### 2017-04-20 Blog/01-NotEquivalent.tex
Adding two blog entries

\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Newtonian, Lagrangian and Hamiltonian mechanics are not equivalent}

After taking a course in advanced mechanics, one is often left with the impression that Netwonian, Lagrangian and Hamiltonian mechanics are all equivalent. Unfortunately, while this is often what a books would say, I don't think that's the case. And you shouldn't take my word for it, or you would make the same mistake I made when I believed everything that was written in my mechanics book.

Textbooks typically stress systems that can be described by all three. We'll go through two counter examples: a body affected by air friction (which is Newtonian but neither Lagrangian or Hamiltonian) and the photon as a particle (which is Hamiltonian but neither Lagrangian or Newtonian).

We are going to be only looking at time independent (i.e. force/Hamiltonian/Lagrangian does not depend on time) particle mechanics (not field theories). In those cases you get slightly different conclusions.

\section{Quick refresh}

Let's go over some key definitions first, so that we have a precise idea of what we are talking about.

A system is Netwonian if there is a mass $m$ and a force $F(x,v)$ such that Newton's second law applies:
\begin{equation}
\label{Fma}
F=ma
\end{equation}

A system is Lagrangian if there is a function $L(x,v)$ such that the system obeys the Euler-Lagrange equations:
\begin{equation}
\label{EulerLagrange}
\frac{\partial L}{\partial x} - \frac{d}{dt} \frac{\partial L}{\partial v} = 0
\end{equation}
Note that, using the chain rule, we have:
\begin{align*}
\frac{d}{dt} \frac{\partial L}{\partial v} &= \frac{\partial}{\partial x} \frac{\partial L}{\partial v} \frac{dx}{dt} + \frac{\partial}{\partial v}  \frac{\partial L}{\partial v} \frac{dv}{dt} \\
&= \frac{\partial^2 L}{\partial x \partial v} v + \frac{\partial^2 L}{\partial^2 v} a
\end{align*}
We can then rewrite the Euler-Lagrange equations as:
\begin{equation}
\label{EulerLagrangeMod}
\frac{\partial^2 L}{\partial^2 v} a + \frac{\partial^2 L}{\partial x \partial v} v - \frac{\partial L}{\partial x}=0
\end{equation}
Which will make our job easier.

A system is Hamiltonian if there is a function $H(x,p)$ such that the system follows Hamilton's equations:
\begin{equation}
\begin{aligned}
\frac{dx}{dt} &= \frac{\partial H}{\partial p} \\
\frac{dp}{dt} &= - \frac{\partial H}{\partial x}
\end{aligned}
\label{Hamilton}
\end{equation}

In many cases, the same system admits all three descriptions. A free particle is such a system, as
\begin{equation}
\begin{aligned}
F&=0 \\
L&=\frac{1}{2}mv^2 \\
H&=\frac{1}{2}\frac{p^2}{m}
\end{aligned}
\end{equation}
yield the same trajectories. The claim is that the following two systems do not.

\section{Massive particle under linear drag}

Suppose we have a massive particle subject to air friction, modeled by linear drag. In this case $F=-bv$ and we have:
\begin{equation}
\label{dragEq}
ma+bv=0
\end{equation}
Is this compatible with Lagrangian and Hamiltonian mechanics?

For \eqref{EulerLagrangeMod} to give us \eqref{dragEq}, we must have
\begin{align}
\label{dragMass}
&\frac{\partial^2 L}{\partial^2 v} = m \\
\label{dragVelocity}
&\frac{\partial^2 L}{\partial x \partial v} v - \frac{\partial L}{\partial x} = bv
\end{align}

Integrating \eqref{dragMass}, the form of the Lagrangian is restricted to:
\begin{equation*}
L = \frac{1}{2} m v^2 + f_1(x) v + f_2 (x)
\end{equation*}
where $f_1(x)$ and $f_2 (x)$ are arbitrary functions of position. We substitute this into \eqref{dragVelocity} and have:
\begin{align*}
\frac{\partial^2 L}{\partial x \partial v} v - \frac{\partial L}{\partial x} &= (\frac{\partial f_1}{\partial x}) v - \frac{\partial f_1}{\partial x} v -  \frac{\partial f_2}{\partial x} \\
&= -  \frac{\partial f_2}{\partial x} \\
&= bv
\end{align*}
which is impossible as $f_2$ is only a function of position. Therefore there is no Lagrangian that would give us \eqref{dragEq} as equations of motion: a massive particle under linear drag is not a Lagrangian system.

We can intuitively understand that the system is not Hamiltonian either: it is dissipative, it does not conserve energy. As Hamiltonian mechanics can only describe systems that conserve energy (i.e. the Hamiltonian), a massive particle under linear drag is not a Hamiltonian system.

Mathematically, Hamiltonian mechanics does not have attractors. That is, it cannot have regions of phase space, like equilibrium point, toward which the motions tends to go. Suppose it had an equilibrium point, then the volume around it would shrink and shrink during the evolution. This violates Louisville's theorem which states that the phase space volume of an Hamiltonian system is conserved. In this specific case, the region with zero velocity is an attractor for the system: whatever velocity you start with, you end up at rest. This means the system is not Hamiltonian.

Writing down the math for the general case is a bit involved, so we just show that if we assume $p=mv$ Hamilton's equations \eqref{Hamilton} cannot give \eqref{dragEq}. We have:
\begin{equation*}
\frac{p}{m}=v= \frac{dx}{dt} = \frac{\partial H}{\partial p}
\end{equation*}
Integrating, the form of the Hamiltonian must be
\begin{equation*}
H=\frac{1}{2}\frac{p^2}{m} + V(x)
\end{equation*}
where $V$ is an arbitrary function of position. Substitute in the second equation:
\begin{equation*}
\frac{dp}{dt}= ma = - \frac{\partial H}{\partial x} = - \frac{\partial V}{\partial x}
\end{equation*}
$V$ is only a function of position, therefore we cannot obtain a force that is a function of velocity.

A massive particle under drag is not a Lagrangian or a Hamiltonian system.

\section{Photon as a free particle}

Suppose we have a photon traveling in space. It's trajectory is a line, traveled at constant speed $c$. It's momentum $p=\frac{\hbar}{\lambda}$ is conserved, where $\lambda$ is the wavelength. The system is not Newtonian: the mass is zero therefore \eqref{Fma} does not apply. Is it Hamiltonian?

The energy of the particle is $E=\hbar \nu$ where $\nu$ is the frequency. The product of the magnitude of the frequency and wavelength is euqal to the speed: $|\nu \lambda| = c$. Therefore we have:
\begin{equation*}
E=\hbar \nu = c \frac{\hbar}{|\lambda|} = c|p| \end{equation*}
This is the energy as a function of position and momentum: it's the Hamiltonian of the system.

Let's verify that Hamilton's equations \eqref{Hamilton} give us the correct equations.
\begin{align*}
H &= c|p| \\
\frac{dx}{dt} &= \frac{\partial H}{\partial p} = c \frac{\partial |p|}{\partial p} = c \frac{|p|}{p} \\
\frac{dp}{dt} &= - \frac{\partial H}{\partial x} = 0
\end{align*}
The momentum cannot be zero, since $H$ is discontinuous there. The velocity is $c$ in the direction of momentum. The momentum is conserved. The system is correctly described by Hamiltonian mechanics.

Since we have the Hamiltonian, we would be tempted to write:
\begin{equation}
\label{Legendre}
L = pv - H
\end{equation}
Unfortunately, this does not work because $L$ cannot be expressed as a function of $x$ and $v$ ($H$ is not hyperregular). Because the velocity is always constant and is only a function of the direction of $p$, the change from momentum to velocity loses information. In other words: position and velocity are not enough to describe the state of the system.

The photon treated as a particles is not a Lagrangian or Newtonian system.

\section{Conclusion}

Since we were able to find counter examples, Netwonian, Lagrangian and Hamiltonian are not equivalent: they describe different types of systems. Newtonian and Lagrangian mechanics are restricted to massive systems, the ones for which position and velocity are enough to identify the state of the system. Lagrangian and Hamiltonian mechanics are restricted to conservative systems, where no energy and no information is exchanged with the outside, where isolation holds and the system is deterministic and reversible.

``But wait!" You may say. ``This was only for the time independent case! In the time dependent case I can write a Hamiltonian that describes dissipative systems! Thus what you said it's wrong and you don't know what you are talking about." That may be true, but the time dependent case needs to be looked at with care as it is easy to get confused. It will be the topic for another time.

\end{document}

### 2017-04-20 Blog/02-HamiltonianDissipative.tex
Adding two blog entries

\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Can Hamiltonian mechanics describe dissipative systems?}

In a previous post we saw how Hamiltonian and Newtonian mechanics are different as the first cannot describe dissipative systems. Yet, if we allow the Hamiltonian to be time dependent, the system would appear to change its energy. Shouldn't that allow us to describe dissipative systems?

In fact, in the literature one can find time dependent Hamiltonians for damped harmonic oscillators. What is going on? Can Hamiltonian mechanics describe dissipative systems or not? If not, what does it mean when the Hamiltonian changes in time?

As we'll see, with time dependent systems it is easy to get confused. Changes of the time variable may introduce effects that are not usually discussed and therefore expected. We'll work on three different systems: particle under air friction (i.e. linear drag), a system whose mass increases in time and a free particle trajectory under a non-linear time transformation. We'll see that all three system have the same trajectories, so they look the same, but only the last two admit Hamiltonian formulation that matches the predictions for momentum and energy.

\section{Quick refresh}

Let's go over some key definitions first. We'll use Newton's second law:
\begin{equation}
\label{Fm_0a}
F=m_0a
\end{equation}
as it applies to a constant mass $m_0$ and a force $F(x,v)$. We reserve the symbol $m$ for the case of variable mass.

We'll use Hamilton's equations:
\begin{equation}
\begin{aligned}
\frac{dx}{dt} &= \frac{\partial H}{\partial p} \\
\frac{dp}{dt} &= - \frac{\partial H}{\partial x}
\end{aligned}
\label{Hamilton}
\end{equation}
where $H(x,p)$ is the Hamiltonian for the system.

\section{A Hamiltonian for particle under linear drag}

Suppose we have a massive particle subject to air friction, modeled by linear drag. In this case $F=-bv$ and we have:
\begin{equation}
\label{dragEq}
m_0a+bv=0
\end{equation}
This equation admits solutions of the type:
\begin{equation}
\begin{aligned}
x &= x_0+v_0\frac{m_0}{b}(1-e^{-\frac{b}{m_0}t}) \\
v &= v_0 e^{-\frac{b}{m_0}t} \\
a &= -\frac{b}{m_0} v_0 e^{-\frac{b}{m_0}t}
\end{aligned}
\label{dragSol}
\end{equation}

Consider now the Hamiltonian:
\begin{equation}
\label{dragHam}
H = \frac{1}{2} \frac{p^2}{m_0} e^{-\frac{b}{m_0}t}
\end{equation}

If we apply \eqref{Hamilton} we have:
\begin{equation*}
\begin{aligned}
\frac{dx}{dt} &= \frac{\partial H}{\partial p} = \frac{p}{m_0} e^{-\frac{b}{m_0}t} \\ 
\frac{dp}{dt} &= - \frac{\partial H}{\partial x} = 0
\end{aligned}
\end{equation*}
Note how the expression for the velocity in \eqref{dragSol} matches the equation above. The solution therefore is
\begin{equation}
\begin{aligned}
p &= p_0 \equiv m_0 v_0 \\
x &= x_0+\frac{p_0}{m_0}\frac{m_0}{b}(1-e^{-\frac{b}{m_0}t})
\end{aligned}
\label{dragHamSol}
\end{equation}
which looks equivalent to the solution for a massive particle under linear drag.

But there is something not quite right. First of all, the momentum is constant. As the mass is constant and the velocity decreases, we would have expected the momentum to be:
\begin{equation}
\label{dragExpMom}
p = m_0 v = m_0 v_0 e^{-\frac{b}{m_0}t}
\end{equation}

For the energy, we would have expected:
\begin{equation*}
E = \frac{1}{2} m_0 v^2 = \frac{1}{2} m_0 v_0^2 e^{-2\frac{b}{m}t}
\end{equation*}
Since $p = p_0 = m_0 v_0$ we have:
\begin{equation*}
E = \frac{1}{2} \frac{p^2}{m_0} e^{-2\frac{b}{m_0}t}
\end{equation*}
which does not match \eqref{dragHam}.

It seems that we have the right kinematics, the right trajectory in space, but it looks like the dynamics, the momentum and the energy, is off. Are we describing the correct system?

\section{Variable mass system}

If we stare at \eqref{dragExpMom} we may get an idea: if the mass were not a constant, but increased in time at the same exact rate the velocity decreased, the momentum would be constant. Maybe the Hamiltonian in \eqref{dragHam} is really  describing such a system?

Suppose we have a massive body. Suppose all around little tiny specs precipitate on the body, making it more massive. Suppose they would come from all directions equally such that the total contribution to the momentum is zero. What would be the equation of motion for such system?

As time passes, the body increases in mass but the momentum is unchanged. We can write:
\begin{equation}
\label{accFma}
\frac{d}{dt}(m v) = 0 = \frac{dm}{dt}v + m \frac{dv}{dt}
\end{equation}
Note already how similar it is similar to \eqref{dragEq}. As the mass changes, we can write it as:
\begin{equation}
m = m_0 \hat{m}(t)
\end{equation}
so that $m_0$ is the initial mass and $\hat{m}(t)$ is the increase factor (i.e. equal to $1$ at $t_0$). We can rewrite the equation as:
\begin{equation*}
\frac{1}{\hat{m}} (\frac{dm}{dt}v + m \frac{dv}{dt}) = \frac{m_0}{m}\frac{dm}{dt}v + m_0 a
\end{equation*}

Suppose that the rate of accretion, as a fraction of the total mass, is constant. We can call $\lambda$ such constant and have:
\begin{equation}
\label{accMass}
\frac{dm}{dt} \frac{1}{m} = \lambda \equiv \frac{b}{m_0}
\end{equation}
We now have the same equation as \eqref{dragEq}.

While the equation is the same, the physical system we are describing is totally different. They both will describe the same trajectory, they are kinematically equivalent, but the first is losing momentum as it slows down while the second is not. So let's see how the dynamical variables compare.

We can solve \eqref{accMass} to find:
\begin{equation}
\label{accMassSol}
m = m_0 e^{\frac{b}{m} t}
\end{equation}
We can combine this with the velocity from \eqref{dragSol} and have:
\begin{equation}
p = m v = m_0 e^{\frac{b}{m} t} v_0 e^{-\frac{b}{m}t} = m_0 v_0 = p_0
\end{equation}
The momentum is now constant and equal to the initial momentum. This is consistent with \eqref{dragHamSol}.

How about the energy? We have:
\begin{equation}
H = \frac{1}{2} \frac{p^2}{m} = \frac{1}{2} \frac{p^2}{m_0 e^{\frac{b}{m} t}} = \frac{1}{2} \frac{p^2}{m_0} e^{-\frac{b}{m} t}
\end{equation}
which is consistent with \eqref{dragHam}.

The predictions, therefore, seems to be more in line with the variable mass system than with the massive particle under drag. We should be inclined to conclude that the Hamiltonian describes the former and not the latter.

\section{Non-linear time transformations}

There is another interested case: one where system is actually closed (i.e. no exchange of matter and energy). Suppose we have a massive particle with no forces acting on it, and let $\hat{t}$ be the time as measured by an inertial observer:
\begin{equation*}
F=m_0\frac{d^2x}{d\hat{t}^2}=0
\end{equation*}
Suppose we change the time variable without changing position:
\begin{equation}
t=t(\hat{t})
\end{equation}
The new variable only depends on the old time variable and the transformation is locally invertible. What happens to the equation of motion?

We have:
\begin{equation*}
m_0\frac{d^2x}{d\hat{t}^2}= m_0 \frac{d}{d\hat{t}} \frac{dx}{d\hat{t}} =0 = m_0 \frac{dt}{d\hat{t}}\frac{d}{dt} (\frac{dt}{d\hat{t}}\frac{dx}{dt})
\end{equation*}
Which can be rewritten as:
\begin{equation}
\begin{aligned}
m &= m_0 \frac{dt}{d\hat{t}} \equiv m_0 \hat{m}(t) \\
\frac{d}{dt} (m\hat{v}) &= 0
\end{aligned}
\label{transMass}
\end{equation}
Note that is formally identical to \eqref{accFma}. What is going on?

As the time transformation is non-linear, the body velocity and acceleration will appear to change. In fact, it may appear to be harder or easier to accelerate it: the effective mass changes in time according to the time transformation. This is just an effect of the new time variable: the mass is not actually increasing.

What time transformation will give us the motion in \eqref{dragSol}? In the inertial frame, the solution will be
\begin{equation}
x=x_0+v_0\hat{t}
\end{equation}
which means
\begin{equation}
\hat{t} = \frac{m}{b}(1-e^{-\frac{b}{m}t})
\end{equation}
Note that the new time variable cannot describe any moment in time after $\hat{t} = \frac{m}{b}$. The body may appear to stop, but it's really the time variable that is converging to a particular instant of time. We can calculate the effective mass combining this with \eqref{transMass}. We have:
\begin{align*}
\hat{t} &= \frac{m}{b}(1-e^{-\frac{b}{m}t}) \\
\hat{t} &- \frac{m}{b} = -\frac{m}{b}e^{-\frac{b}{m}t} \\
(1&- \frac{b}{m}\hat{t}) = e^{-\frac{b}{m}t} \\
\log(1&- \frac{b}{m}\hat{t})= -\frac{b}{m}t \\
t&= -\frac{m}{b}\log(1- \frac{b}{m}\hat{t}) \\
\frac{dt}{d\hat{t}} &= -\frac{m}{b} \frac{1}{1- \frac{b}{m}\hat{t}} (-\frac{m}{b})  \\
&= \frac{1}{1- \frac{b}{m}\hat{t}}
= \frac{1}{1- \frac{b}{m}\frac{m}{b}(1-e^{-\frac{b}{m}t})} \\
&=\frac{1}{1-1-e^{-\frac{b}{m}t}} =e^{\frac{b}{m}t} \\
m&=m_0\frac{dt}{d\hat{t}}=m_0e^{\frac{b}{m}t}
\end{align*}
which is the same expression as \eqref{accMassSol}

In other words, the two cases of non-linear time transformation and variable mass are indistinguishable: they share the same kinematics and dynamics. In reality, they are as we could look at the motion of other objects around us: that would tell us if we are in an inertial frame or not.

\section{Conclusion}

So, can Hamiltonian mechanics describe dissipative systems? What does the Hamiltonian \eqref{dragHam} describe?

While the kinematics, the position and velocity, agrees with a particle under linear drag, the dynamics, the momentum and energy, does not. I would be inclined to rule it out.

The two other system match in both kinematics and dynamics. The key is they still conserve momentum, even its definition changes in time (i.e. the effective mass is changing). While the energy is changing in time, there is still something conserved by the system.

Between these two, I would be inclined to consider the last one, the non-linear time transformation, to hold physical relevance. The variable mass is just constructed to look the same, but would not hold in the relativistic case: at that point gravity allows to tell the difference.

And the physical insight is the following: a time dependent Hamiltonian system does not allow us to describe an arbitrary system that is changing energy in time. Instead, it provides the correct coordinate independent generalization of non-dissipative systems.

Energy is not invariant under time transformations. From special relativity we know that it is the time component of a covector. Therefore, under time transformation, it changes like:
\begin{equation}
H=\frac{d\hat{t}}{dt}\hat{H}
\end{equation}
Note that this is exactly what the Hamiltonian \eqref{dragHam} is: the free particle Hamiltonian multiplied by the time derivative.

There is a way to rewrite Hamilton's equation in a different but equivalent form, that is manifestly relativistic and introduces an invariant Hamiltonian. But that is a topic for another time.

\end{document}

### 2017-04-26 Blog/03-AllLagrangiansMustBeConvex.tex
Adding new post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Lagrangians must be convex}

TL;DR Only convex Lagrangians can give unique solutions when the action is minimized.

This is the kind of post I do mainly for my future self. Sometimes it takes far longer than I would have expected to nail down a detail and I want a record of the solution so that I can take it out of my head but still be able to look at it later.

For a long time, I couldn't quite understand whether Lagrangians must be convex in velocity (or hyper-regular depending on your mathematical training) or it just happened that the ones we are interested in are. Physics textbook always assume, one way or the other, that the Lagrangian is of a particular form which is indeed convex: kinetic energy minus potential energy. But they don't seem to investigate the more general questions. Are all Lagrangians convex? If so, why? What happens if they are not?

The answer is clear cut, as it usually is once one figures out the details. The Lagrangian must be convex or the problem of minimizing the action is not well posed. To understand the gist, let's see what happens when you want to find the minimum of a differentiable function $f(x)$. You first look for stationary points, where $\dfrac{df}{dx} = 0$. Not all these are minimums: they could maximums or inflection points as well. So you look at the second derivative. If it is positive, $\dfrac{d^2f}{dx^2} > 0$, then it is a maximum. If it's zero, you continue until either you get a non-zero odd derivative (not a maximum) or a positive even derivative (a maximum).

If it's a function of multiple variables, the first derivative is replaced by the gradient $\nabla f \equiv \dfrac{\partial f}{\partial x^i} = 0$; the second derivative is replaced by the Hessian, $\textrm{Hess}(f) \equiv \nabla \nabla f = \left[\frac{d^2f}{dx^i dx^j}\right]$, which must be positive definite, $\frac{d^2f}{dx^i dx^j} x^i x^j > 0$, meaning that the second derivative along any direction is positive and we have a minimum along any direction (i.e. all eigenvalues positive).

To minimize the action, something similar needs to happen. We first want the action to be stationary, $\delta S = 0$. But this is not sufficient: we need to check that we are in an actual maximum. To do that, we look at the second variation and see whether it is positive, $\delta S > 0$. We'll see that this leads to the requirement $\frac{d^2L}{dv^i dv^j} v^i v^j > 0$. Which means the matrix
\begin{equation}
\begin{bmatrix}
\dfrac{\partial^2 L}{\partial v_1^2} & \dfrac{\partial^2 L}{\partial v_1\,\partial v_2} & \cdots & \dfrac{\partial^2 L}{\partial v_1\,\partial v_n} \\[2.2ex]
\dfrac{\partial^2 L}{\partial v_2\,\partial v_1} & \dfrac{\partial^2 L}{\partial v_2^2} & \cdots & \dfrac{\partial^2 L}{\partial v_2\,\partial v_n} \\[2.2ex]
\vdots & \vdots & \ddots & \vdots \\[2.2ex]
\dfrac{\partial^2 L}{\partial v_n\,\partial v_1} & \dfrac{\partial^2 L}{\partial v_n\,\partial v_2} & \cdots & \dfrac{\partial^2 L}{\partial v_n^2}
\end{bmatrix}
\label{lagHessian}
\end{equation}
is positive definite everywhere and the Lagrangian is convex.

The reason that only the mixed derivatives with respect to velocity appear is that one can always create arbitrary small variations in position with arbitrarily large variations in velocity. As such, the Lagrangian needs to be convex in velocity or we can always create a variation that reduces the action.

But before we begin, a rant if I may. Textbooks seem to skimp on these details and most seem to cite results from Galfand and Fomin ``Calculus of Variation" and suggest to look there for details. %http://users.uoa.gr/~pjioannou/mech2/READING/Gelfand_Fomin_Calculus_of_Variations.pdf
The single degree of freedom was straightforward enough to figure out by myself, but I didn't seem to be able to generalize it to the multidimensional case mainly because $\dfrac{d^2L}{dx^i dv^j}$ is not symmetric. To my surprise, Galfand and Fomin used the same trick I used and simply assumed $\dfrac{d^2L}{dx^i dv^j}$ to be symmetric. And in a footnote they add ``Without this assumption, which is unnecessarily restrictive, equations (54) and (55) become more complicated, but it can be shown that Theorems 1 and 2 remain valid (H. Niemeyer, private communication)." Well, no sh*t Sherlock! The assumption \textbf{is} too restrictive and the math \textbf{does} get more complicated: that's why I couldn't figure it out! So, yeah: all the textbooks that refer to Galfand and Fomin for this they are referring to their private communication! Anyway, I was able to find a way that is simple enough and this is what I present here.

\section{Lagrangians must be convex}

As usual, let's start with a Lagrangian, this time defined on multiple dimensions.
\begin{equation}
L(t, x^i, v^i)
\end{equation}
The action is the integral of the Lagrangian over a path
\begin{equation}
S = \int_{t_0}^{t_1} L(t, x^i(t), v^i(t)) dt \\
\end{equation}
We set the action to be stationary
\begin{equation}
\delta S = 0
\end{equation}
and get the Euler-Lagrange equations
\begin{equation}
\label{EulerLagrange}
\frac{\partial L}{\partial x^i} - \frac{d}{dt} \frac{\partial L}{\partial v^i} = 0
\end{equation}
Using the chain rule, we have:
\begin{align*}
\frac{d}{dt} \frac{\partial L}{\partial v^i} &= \frac{\partial}{\partial x^j} \frac{\partial L}{\partial v^i} \frac{dx^j}{dt} + \frac{\partial}{\partial v^j}  \frac{\partial L}{\partial v^i} \frac{dv^j}{dt} \\
&= \frac{\partial^2 L}{\partial v^i \partial x^j} v^j + \frac{\partial^2 L}{\partial v^i \partial v^j} a^j
\end{align*}
We can then rewrite the Euler-Lagrange equations as:
\begin{equation}
\label{EulerLagrangeMod}
\frac{\partial^2 L}{\partial v^i \partial v^j} a^j + \frac{\partial^2 L}{\partial v^i \partial x^i} v^j - \frac{\partial L}{\partial x^i}=0
\end{equation}

The first element in the equation is the matrix we saw in \eqref{lagHessian}. For the equation to have a unique solution, we must be able to write the acceleration in terms of position and velocity. That is, we need $\dfrac{\partial^2 L}{\partial v^i \partial v^j}$ to be invertible, its determinant has to be non-zero:
\begin{equation}
\label{lagHessianNonSingular}
\left|\frac{\partial^2 L}{\partial v^i \partial v^j}\right| \neq 0
\end{equation}

To be sure that the action at a minimum (e.g. not maximum or at a saddle point) we must make sure that variations can only produce greater values. We want to make sure, then, that the second variation cannot create a negative change. That is:
\begin{equation}
\delta^2 S \geq 0
\end{equation}

We have:
\begin{align}
\delta^2 S &= \int_{t_0}^{t_1} \delta^2 L(t, x^i(t), v^i(t)) dt = \int_{t_0}^{t_1} \delta \left[\frac{\partial L}{\partial x^i} \delta x^i +  \frac{\partial L}{\partial v^i} \delta v^i \right]dt \nonumber \\
&= \int_{t_0}^{t_1} \left[ \frac{\partial^2 L}{\partial x^i\partial x^j} \delta x^i \delta x^j + \frac{\partial^2 L}{\partial x^i\partial v^j} \delta x^i \delta v^j \right. + \left. \frac{\partial^2 L}{\partial v^i\partial x^j} \delta v^i \delta x^j  + \frac{\partial^2 L}{\partial v^i\partial v^j} \delta v^i \delta v^j\right]dt
\label{secvar}
\end{align}

This has to be greater or equal to zero for any variation. In particular we consider this family of variations: oscillation of $\hat{n}$ cycles between two instants $\hat{t}_0$ and $\hat{t}_1$ along the direction given by the vector with components $\hat{x}^i$. For example, $3$ oscillation from $1.2s$ and $1.3s$ along the $x^1$ direction. The variation can be written as:
\begin{equation}
\label{posvar}
\delta x^i = \left\{
\begin{array}{ll}
\hat{x}^i \sin\left(\frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} (t-\hat{t}_0)\right)  & \hat{t}_0 \leq t \leq \hat{t}_1 \\
0 & t < \hat{t}_0 \; \textrm{or} \; t > \hat{t}_1
\end{array}
\right.
\end{equation}
Note that the change for the velocity:
\begin{equation}
\label{velvar}
\delta v^i = \left\{
\begin{array}{ll}
\hat{x}^i \frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} \cos\left(\frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} (t-\hat{t}_0)\right)  & \hat{t}_0 \leq t \leq \hat{t}_1 \\
0 & t < \hat{t}_0 \; \textrm{or} \; t > \hat{t}_1
\end{array}
\right.
\end{equation}
depends linearly on $\hat{n}$. As we increase the number of oscillations in the same interval, the velocity at which those oscillations are preformed increases. That is: even if the change in position remains bounded, the change in velocity can be made arbitrarily large.  If we substitute \eqref{posvar} and \eqref{velvar} into \eqref{secvar} we have:
\begin{align*}
\delta^2 S &= \int_{\hat{t}_0}^{\hat{t}_1} \left[ \frac{\partial^2 L}{\partial x^i\partial x^j} \hat{x}^i \hat{x}^j \sin^2\left(\frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} (t-\hat{t}_0)\right) \right. \\
& + \left. 2 \frac{\partial^2 L}{\partial x^i\partial v^j} \hat{x}^i \hat{x}^j \frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} \sin\left(\frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} (t-\hat{t}_0)\right) \cos\left(\frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} (t-\hat{t}_0)\right) \right. \\
& + \left.  \frac{\partial^2 L}{\partial v^i\partial v^j} \hat{x}^i \hat{x}^j \frac{4\pi^2 \hat{n}^2}{(\hat{t}_1 - \hat{t}_0)^2} \cos^2\left(\frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} (t-\hat{t}_0)\right) \right]dt > 0
\end{align*}

The integral is now restricted only where the variation is non-zero. As $\hat{n}$ increases, the third term is the one that dominates: it increases as $\hat{n}^2$. The third term, then, cannot be negative as the variation would net a negative contribution to the action: the path would not be minimizing the action. We have:
\begin{equation}
\int_{\hat{t}_0}^{\hat{t}_1} \frac{\partial^2 L}{\partial v^i\partial v^j} \hat{x}^i \hat{x}^j \frac{4\pi^2 \hat{n}^2}{(\hat{t}_1 - \hat{t}_0)^2} \cos^2\left(\frac{2\pi \hat{n}}{\hat{t}_1 - \hat{t}_0} (t-\hat{t}_0)\right) dt \geq 0
\end{equation}
The inequality has to hold for arbitrary choices of $\hat{t}_0$ and $\hat{t}_1$, which means the integrand has to be non-negative at every $t$. Also note that the right part of the integrand is already non-negative, therefore we have:
\begin{equation}
\label{lagHessianNonNegative}
\frac{\partial^2 L}{\partial v^i\partial v^j} \hat{x}^i \hat{x}^j \geq 0
\end{equation}

If we combine \eqref{lagHessianNonSingular} with \eqref{lagHessianNonNegative} we get that
\begin{equation}
\label{lagHessianPositive}
\frac{\partial^2 L}{\partial v^i\partial v^j} \hat{x}^i \hat{x}^j > 0
\end{equation}
is positive definite. Since we assumed the Lagrangian to be twice differentiable, it means it will be convex.

\section{Conclusion}

As we have seen, the convexity of the Lagrangian is a constraint that is necessary for the existence of the solution of the minimization problem. Essentially: a non convex Lagrangian is not a well posed problem. Also note that this is just a necessary condition: it is not sufficient.

Consider a free particle on a sphere. If you take to points, the trajectory that minimizes the action is the shortest arc traveled at constant speed. Now take two opposite points: all arcs are of the same length, all of them will give the same action. There is no single minimum. The convexity of the Lagrangian is a local condition. One still needs to make sure that the global problem is solvable.

But it is enough to make the Euler-Lagrange equations solvable. That is: given initial position and velocity on a sphere, we can calculate the trajectory. For this reason, I am not convinced the global minimization problem is of physical significance. I do think the local minimization problem is, though. But this is a topic for another time.

\end{document}

### 2017-04-26 Blog/02-HamiltonianDissipative.tex
Minor corrections

@@
-\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
+\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}
 
@@
 
+TL;DR Time dependent Hamiltonian mechanics is for conservative forces in non inertial frames, not for non-conservative forces.
+
 In a previous post we saw how Hamiltonian and Newtonian mechanics are different as the first cannot describe dissipative systems.
@@
 We'll see that all three system have the same trajectories, so they look the same, but only the last two admit Hamiltonian formulation that matches the predictions for momentum and energy.
+In both those cases inertia is conserved but the mass changes, or appears to change in time, making the Hamiltonian time dependent.
 
@@
 
-As the time transformation is non-linear, the body velocity and acceleration will appear to change.
+As the time transformation is non-linear, the body's velocity and acceleration will appear to change.
 In fact, it may appear to be harder or easier to accelerate it: the effective mass changes in time according to the time transformation.
@@
 In other words, the two cases of non-linear time transformation and variable mass are indistinguishable: they share the same kinematics and dynamics.
-In reality, they are as we could look at the motion of other objects around us: that would tell us if we are in an inertial frame or not.
+Well, we could distinguish between them by looking at the motion of other objects around us: that would tell us if we are in an inertial frame or not.
 
@@
 
-Between these two, I would be inclined to consider the last one, the non-linear time transformation, to hold physical relevance.
-The variable mass is just constructed to look the same, but would not hold in the relativistic case: at that point gravity allows to tell the difference.
+Note that if we added gravity (i.e.
+general relativity), we might be able to notice that at some point the exponentially increasing mass turns into a black hole.
+The third case, though, would remain the same.
+Therefore, I would be inclined to consider the last one, the non-linear time transformation, to hold the most physical relevance.
 
-And the physical insight is the following: a time dependent Hamiltonian system does not allow us to describe an arbitrary system that is changing energy in time.
-Instead, it provides the correct coordinate independent generalization of non-dissipative systems.
-
-Energy is not invariant under time transformations. From special relativity we know that it is the time component of a covector. Therefore, under time transformation, it changes like:
+The physical insight is the following: a time dependent Hamiltonian system does not allow us to describe an arbitrary system that is changing energy in time.
+Only Newtonian mechanics can do that.
+Instead, it provides the correct coordinate independent generalization for conservative forces.
+Energy is not invariant under time transformations.
+From special relativity we know that it is the time component of a covector.
+Therefore, under time transformation, it changes like:
 \begin{equation}

### 2017-05-09 Blog/04-LagrangianSubsetOfHamiltonian.tex
Post on relationship between Lagrangian and Hamiltonian mechanics

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Lagrangian mechanics is a subset of Hamiltonian mechanics}

TL;DR Lagrangians must be convex while Hamiltonians may not. Lagrangian systems are Hamiltonian systems where the momentum is a function $p=p(x,v)$ of position and velocity, strictly increasing with respect to velocity.

In a previous post we saw how the photon, as a particle, is a Hamiltonian system but not a Lagrangian system, which shows how the two concepts are not equivalent. So, what is the relationship between Hamiltonian and Lagrangian mechanics? How do they overlap?

We'll see that all the Lagrangians that lead to solvable equations of motion admit a Hamiltonian (i.e. they must be convex/hyper-regular) but not all Hamiltonians that lead to solvable equations of motions admit a Lagrangian (i.e. they may not be convex/hyper-regular). This means that every Lagrangian system is also a Hamiltonian system, but not vice-versa.

\section{Lagrangians systems are Hamiltonian}

Let's refresh our math. We have a Lagrangian
\begin{equation}
\label{Lagrangian}
L(x^i,v^i)
\end{equation} 
function of position and velocity. We define conjugate momentum as
\begin{equation}
\label{MomentumFromLagrangian}
p_i = \frac{\partial L}{\partial v^i} = p_i(x^j,v^j)
\end{equation}
where each component is a function of position and velocity. We define the Hamiltonian as
\begin{equation}
\label{HamiltonianFromLagrangian}
H = v^i p_i - L = H(x^j, p_j)
\end{equation}
expressed as a function of position and momentum. One can then show that the Euler-Lagrange equations are equivalent to Hamilton equations.

Now the question is: when can we actually do this? When are we allowed to define a Hamiltonian for a Lagrangian?

First of all, the derivatives of the Lagrangian with respect to velocity must exist and be non-zero or we can't even define momentum. Not only that, we need to be able to relate changes in momentum to the ones in position and velocity: the transformation must be differentiable, the Lagrangian is twice differentiable. Then we must be able to express the Hamiltonian as a function of position and momentum. To be able to do that, we must be able to write the velocity in terms of position and momentum: \eqref{MomentumFromLagrangian} must be invertible at every point. This means that the  Jacobian $\left| \dfrac{\partial p_i}{\partial v^j} \right| = \left| \dfrac{\partial^2 L}{\partial v^i \partial v^j} \right| \neq 0$, which is the Hessian of the Lagrangian over velocity, must be non-singular.

But for the minimization of the action to be locally well defined, as we saw in a previous post, the Lagrangian must be convex in $v^i$. This means that the Jacobian of the transformation between velocity and momentum is positive at every point: we can write the Hamiltonian. So each twice differentiable Lagrangian that admits unique solutions also admits a corresponding Hamiltonian. Each Lagrangian system is also a Hamiltonian system.

The fact that the Lagrangian is convex will also mean that the Hamiltonian will be convex as well: this is a property of the Legendre transformation \eqref{HamiltonianFromLagrangian} that provides the link between them. So the question is: are all Hamiltonian convex?

\section{Not all Hamiltonian systems are Lagrangian}

Let's refresh our math again. We have a Hamiltonian
\begin{equation}
\label{Hamiltonian}
H(x^i, p_i)
\end{equation}
and the equations of motion
\begin{equation}
\begin{aligned}
\frac{dx}{dt} &= \frac{\partial H}{\partial p} \\
\frac{dp}{dt} &= - \frac{\partial H}{\partial x}
\end{aligned}
\label{HamiltonEquations}
\end{equation}
We assume the Hamiltonian to be twice differentiable: this allows phase space volumes to remain well defined during the evolution and, therefore, densities over phase space.

Under what conditions does this set of equations have a unique solution?

This is a set of ordinary differential equations so the Picard Lindelöf theorem applies: if the right side is Lipschitz continuous then the solution exists and it is unique. But the right side are just the derivatives of $H$, which are differentiable because we assumed $H$ is twice differentiable. All differentiable functions are Lipschitz continuous therefore if $H$ is twice differentiable the initial value problem is always defined.

This means: the Hamiltonian does not have to be convex. In fact, we can set the Hamiltonian to be zero (i.e. $H=0$): we simply get $x(t)=x_0$ and $p(t)=p_0$. Contrast this with setting the Lagrangian to zero. The action for all paths would be equal to $0$, all paths minimize it: there is no unique solution.

\section{Conclusion}

All Lagrangian systems can be reformulated as Hamiltonian systems while not all Hamiltonian systems can be reformulated as Lagrangian systems. This means that Lagrangian systems are a subset of Hamiltonian system. But which one exactly?

It's the one in which the velocity can be expressed as an invertible function of position and momentum (we can always calculate one form the other) and the function is strictly increasing in momentum (a positive increase in momentum always yields a positive increase in velocity and vice-versa). This means that a positive volume in momentum is a positive volume in velocity: positive densities in position and momentum are positive densities in position and velocity. If these conditions are met, the Hamiltonian is convex in momentum and admits a Lagrangian.

An intuitive way to remember this: Lagrangian systems are those Hamiltonian system for which each state can be recognized just by looking at its trajectory. If two states have the same trajectory (e.g. the velocity of a photon is always the same no matter the magnitude of momentum) then the system is not Lagrangian.

\end{document}

### 2017-05-23 Blog/03-AllLagrangiansMustBeConvex.tex
Editing from Christine

@@
 For a long time, I couldn't quite understand whether Lagrangians must be convex in velocity (or hyper-regular depending on your mathematical training) or it just happened that the ones we are interested in are.
-Physics textbook always assume, one way or the other, that the Lagrangian is of a particular form which is indeed convex: kinetic energy minus potential energy.
+Physics textbooks always assume, one way or the other, that the Lagrangian is of a particular form which is indeed convex: kinetic energy minus potential energy.
 But they don't seem to investigate the more general questions.
@@
 You first look for stationary points, where $\dfrac{df}{dx} = 0$.
-Not all these are minimums: they could maximums or inflection points as well.
+Not all these are minima: they could be maxima or inflection points as well.
 So you look at the second derivative.
-If it is positive, $\dfrac{d^2f}{dx^2} > 0$, then it is a maximum.
-If it's zero, you continue until either you get a non-zero odd derivative (not a maximum) or a positive even derivative (a maximum).
+If it is positive, $\dfrac{d^2f}{dx^2} > 0$, then it is a minimum.
+If it's zero, you continue until either you get a non-zero odd derivative (not a minimum) or a positive even derivative (a minimum).
 
-If it's a function of multiple variables, the first derivative is replaced by the gradient $\nabla f \equiv \dfrac{\partial f}{\partial x^i} = 0$; the second derivative is replaced by the Hessian, $\textrm{Hess}(f) \equiv \nabla \nabla f = \left[\frac{d^2f}{dx^i dx^j}\right]$, which must be positive definite, $\frac{d^2f}{dx^i dx^j} x^i x^j > 0$, meaning that the second derivative along any direction is positive and we have a minimum along any direction (i.e.
+If it's a function of multiple variables, the first derivative is replaced by the gradient $\nabla f \equiv \dfrac{\partial f}{\partial x^i} = 0$; the second derivative is replaced by the Hessian, $\textrm{Hess}(f) \equiv \nabla \nabla f = \left[\frac{\partial^2f}{\partial x^i \partial x^j}\right]$, which must be positive definite, $\frac{\partial^2f}{\partial x^i \partial x^j} x^i x^j > 0$, meaning that the second derivative along any direction is positive and we have a minimum along any direction (i.e.
 all eigenvalues positive).
@@
 We first want the action to be stationary, $\delta S = 0$.
-But this is not sufficient: we need to check that we are in an actual maximum.
-To do that, we look at the second variation and see whether it is positive, $\delta S > 0$.
-We'll see that this leads to the requirement $\frac{d^2L}{dv^i dv^j} v^i v^j > 0$.
+But this is not sufficient: we need to check that we are in an actual minimum.
+To do that, we look at the second variation and see whether it is positive, $\delta^2 S > 0$.
+We'll see that this leads to the requirement $\frac{\partial^2L}{\partial v^i \partial v^j} v^i v^j > 0$.
 Which means the matrix
@@
 
-The reason that only the mixed derivatives with respect to velocity appear is that one can always create arbitrary small variations in position with arbitrarily large variations in velocity.
+The reason that only the mixed derivatives with respect to velocity appear is that one can always create arbitrarily small variations in position with arbitrarily large variations in velocity.
 As such, the Lagrangian needs to be convex in velocity or we can always create a variation that reduces the action.
@@
 But before we begin, a rant if I may.
-Textbooks seem to skimp on these details and most seem to cite results from Galfand and Fomin ``Calculus of Variation" and suggest to look there for details.
+Textbooks seem to skimp on these details and most seem to cite results from Galfand and Fomin ``Calculus of Variations" and suggest to look there for details.
 %http://users.uoa.gr/~pjioannou/mech2/READING/Gelfand_Fomin_Calculus_of_Variations.pdf
-The single degree of freedom was straightforward enough to figure out by myself, but I didn't seem to be able to generalize it to the multidimensional case mainly because $\dfrac{d^2L}{dx^i dv^j}$ is not symmetric.
-To my surprise, Galfand and Fomin used the same trick I used and simply assumed $\dfrac{d^2L}{dx^i dv^j}$ to be symmetric.
-And in a footnote they add ``Without this assumption, which is unnecessarily restrictive, equations (54) and (55) become more complicated, but it can be shown that Theorems 1 and 2 remain valid (H.
+The single degree of freedom was straightforward enough to figure out by myself, but I didn't seem to be able to generalize it to the multidimensional case mainly because $\dfrac{\partial^2L}{\partial x^i \partial v^j}$ is not symmetric.
+To my surprise, Galfand and Fomin used the same trick I used and simply assumed $\dfrac{\partial^2L}{\partial x^i \partial v^j}$ to be symmetric.
+And in a footnote they add, ``Without this assumption, which is unnecessarily restrictive, equations (54) and (55) become more complicated, but it can be shown that Theorems 1 and 2 remain valid (H.
 Niemeyer, private communication)." Well, no sh*t Sherlock!
 The assumption \textbf{is} too restrictive and the math \textbf{does} get more complicated: that's why I couldn't figure it out!
-So, yeah: all the textbooks that refer to Galfand and Fomin for this they are referring to their private communication!
+So, yeah: all the textbooks that refer to Galfand and Fomin for this are referring to their private communication!
 Anyway, I was able to find a way that is simple enough and this is what I present here.
@@
 
-To be sure that the action at a minimum (e.g.
-not maximum or at a saddle point) we must make sure that variations can only produce greater values.
+To be sure that the action is at a minimum (e.g.
+not a maximum or at a saddle point) we must make sure that variations can only produce greater values.
 We want to make sure, then, that the second variation cannot create a negative change.
@@
 In particular we consider this family of variations: oscillation of $\hat{n}$ cycles between two instants $\hat{t}_0$ and $\hat{t}_1$ along the direction given by the vector with components $\hat{x}^i$.
-For example, $3$ oscillation from $1.2s$ and $1.3s$ along the $x^1$ direction.
+For example, $3$ oscillations from $1.2$ s and $1.3$ s along the $x^1$ direction.
 The variation can be written as:
@@
 depends linearly on $\hat{n}$.
-As we increase the number of oscillations in the same interval, the velocity at which those oscillations are preformed increases.
+As we increase the number of oscillations in the same interval, the velocity at which those oscillations are performed increases.
 That is: even if the change in position remains bounded, the change in velocity can be made arbitrarily large.
@@
 As $\hat{n}$ increases, the third term is the one that dominates: it increases as $\hat{n}^2$.
-The third term, then, cannot be negative as the variation would net a negative contribution to the action: the path would not be minimizing the action.
+The third term, then, cannot be negative as the variation would get a negative contribution to the action: the path would not be minimizing the action.
 We have:
@@
 As we have seen, the convexity of the Lagrangian is a constraint that is necessary for the existence of the solution of the minimization problem.
-Essentially: a non convex Lagrangian is not a well posed problem.
+Essentially: a non-convex Lagrangian is not a well posed problem.
 Also note that this is just a necessary condition: it is not sufficient.
@@
 Consider a free particle on a sphere.
-If you take to points, the trajectory that minimizes the action is the shortest arc traveled at constant speed.
+If you take two points, the trajectory that minimizes the action is the shortest arc traveled at constant speed.
 Now take two opposite points: all arcs are of the same length, all of them will give the same action.

### 2017-05-23 Blog/05-ClassicalMechanicsInOnePost.tex
Adding new post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Classical mechanics in one post}

TL;DR Classical mechanics describes a divergence free flow of states. The Hamiltonian is the time component of its vector potential. The Lagrangian is the scalar product between the flow and the vector potential.

For the longest time I didn't know what the Lagrangian was and why its integral is minimized along trajectories. A couple of months ago I went through TODO and found some insights in differential geometry that got me going and finally worked out a satisfactory answer. I unpacked that understanding in a way that only needs differential calculus, so that it's more understandable and gives direct physics intuition. If you have studied magnetostatic (i.e. the magnetic field and its vector potential) you should be able to fully follow this.

\section{Deterministic and reversible flow}

Take the space formed by position $x$, momentum $p$ and time $t$. This is sometimes called the extended phase space: it represents all states at all times.

The evolution of a state in time is represented by a line in this space. If we have a law of evolution, at each point we have a vector that tells us where the state is going to move. This forms a vector field
\begin{equation}
\begin{aligned}
\vec{S} &= \dfrac{dx}{dt} e_x + \dfrac{dp}{dt} e_p + \dfrac{dt}{dt} e_t \\ 
&= S^x e_x + S^p e_p + S^t e_t 
\end{aligned}
\label{flow}
\end{equation}
 on the extended phase space, which we call flow. $e_x$, $e_p$ and $e_t$ are the unit vectors along each direction, which are multiplied by the respective components of $\vec{S}$.

Note that, since we are using time both as a coordinate and as the parametrization of the curve, the time component of $\vec{S}$ is unitary: $S^t = \dfrac{dt}{dt} = 1$. Also, we are going to assume that the surface integral of $\vec{S}$ on a closed curve is zero: the density of states that flows in is equal to the density of states that flows out. In this sense, I call this deterministic and reversible flow. Well, there are a dozen of other reasons to call it that, but we will see them some other time.

If you remember your vector calculus, you know that we can define a vector potential for a divergence free field. We therefore define
\begin{equation}
\begin{aligned}
\vec{\theta} &= p e_x + 0 e_p - H(x,p,t) e_t \\ 
&= \theta^x e_x + \theta^p e_p + \theta^t e_t 
\end{aligned}
\label{flow}
\end{equation}
This is actually the opposite of vector potential: we need an annoying minus sign to match with existing conventions. We have
\begin{equation}
\begin{aligned}
\vec{S} &= \mathrm{curl}(-\theta) = - \nabla \times \vec{\theta} \\
S^x &= \dfrac{dx}{dt} = - \partial_p \theta^t + \partial_t \theta^p = \partial_p H + \partial_t 0 = \partial_p H \\
S^p &= \dfrac{dp}{dt} = - \partial_t \theta^q + \partial_q \theta^t = \partial_t p - \partial_q H = - \partial_q H \\
S^t &= \dfrac{dt}{dt} = - \partial_q \theta^p + \partial_p \theta^q = - \partial_q 0 + \partial_p p = 1
\end{aligned}
\label{HamiltonEquations}
\end{equation}

We recognize Hamilton's equations. Geometrically, the Hamiltonian is the time component of the vector potential while conjugate momentum is the spatial component.

\section{Line integral of the vector potential}

The line integral of a vector field $\vec{\theta}$ has an interesting, yet little known property: it is stationary stationary along the field lines $\gamma$ of the curl. That is: if we make a small change in the line $\gamma$, the line integral $\int \vec{\theta} \cdot d\vec{\gamma}$ does not change. In other words:

\begin{equation}
\delta \int \vec{\theta} \cdot d\vec{\gamma} = 0
\label{stationaryLineIntegral}
\end{equation}

\begin{figure}
\begin{tikzpicture}

\draw  plot[smooth, tension=.7] coordinates {(0.5,3.5) (0,2.5) (-0.5,0) (-1.5,-1.5) (-2.5,-2.5)};

\node at (-0.75,-0.1) {$\gamma$};

\draw  plot[smooth, tension=.7] coordinates {(0,2.5) (0,1.5) (0.25,0.5) (0,-0.5) (-1,-1.1) (-1.5,-1.5)};

\node at (0.25,-0.75) {$\gamma$'};

\draw[style=ultra thin]  plot[smooth, tension=.7] coordinates {(-0.7,-0.8) (-0.6,-0.9) (-0.6,-1.1) (-0.5,-1.2)};

\node at (-0.3,-1.3) {$\Sigma$};

%\node at (0,0) (Circle) {};
%\draw [<-] (Circle.south)arc(160:-160:-0.2);

\draw [-latex](-0.5,2.5) -- (-0.4,2.9);
\draw [-latex](-1.5,2.5) -- (-1.45,2.9);
\draw [-latex](-2.5,2.5) -- (-2.5,2.9);

\draw [-latex](-0.5,1.5) -- (-0.45,1.9);
\draw [-latex](-1.5,1.5) -- (-1.45,1.9);
\draw [-latex](-2.5,1.5) -- (-2.45,1.9);

\draw [-latex](-0.5,0.5) -- (-0.4,0.9);
\draw [-latex](-1.5,0.5) -- (-1.4,0.9);
\draw [-latex](-2.5,0.5) -- (-2.4,0.9);

\draw [-latex](-1.5,-0.5) -- (-1.3,-0.1);
\draw [-latex](-2.5,-0.5) -- (-2.4,-0.1);

\draw [-latex](-1.5,-1.5) -- (-1.2,-1.1);
\draw [-latex](-2.5,-1.5) -- (-2.2,-1.1);

\node at (-1.5, 1.25) {$\vec{S} = -\,$curl $\vec{\theta}$};

\end{tikzpicture}
\end{figure}

A way to intuitively understand this is the following. Take a field line $\gamma$, a line that is always tangent to the vector field $\vec{S}$. Between two point, make a small change to it $\gamma'$. Take $\Sigma$ to be a surface enclosed by these lines. The line integral of $-\vec{\theta}$ over its contour is given by the flow of its curl $\vec{S}$ through $\Sigma$. If the variation is small enough, the surface $\Sigma$ is tangent to the vector field $\vec{S}$ since $\gamma$ is always tangent to it. The line integral of $-\vec{\theta}$ over its contour is zero. This means that the line integral of $-\vec{\theta}$ along $\gamma$ is equal to the one along $\gamma'$: small variations of $\gamma$ do not change the line integral.

As $-\vec{\theta}$ is the vector potential for the flow $\vec{S}$, we can write
\begin{equation}
\begin{aligned}
\delta \int \vec{\theta} \cdot d\vec{\gamma} &= \delta \int \vec{\theta} \cdot \vec{S} dt \\
&= \delta \int (\theta^x S^x + \theta ^p S^p + \theta^t S^t)dt \\
&= \delta \int \left( p \frac{dx}{dt} + 0 \frac{dp}{dt} - H \frac{dt}{dt} \right)dt \\
&= \delta \int \left( p \frac{dx}{dt} - H  \right)dt \\
&= \delta \int L dt =  0
\end{aligned}
\label{stationaryAction}
\end{equation}
where $L= \vec{\theta} \cdot \vec{S}$ is the Lagrangian. We recognize the principle of stationary action that gives us the Euler-Lagrange equations. Geometrically, the Lagrangian is the scalar product between the flow and (the opposite of) its vector potential.

\section{Conclusion}

I hope you can appreciate how we have condensed the geometrical description of classical mechanics, which usually takes hundreds of pages to explain, in two pages of simple vector calculus. Once you get used to thinking in these terms, you realize that there are not many moving parts.

This clarifies another important aspect: the conjugate momentum $p$, the Hamiltonian $H$ and the Lagrangian $L$ are all ``unphysical" in the sense that they depend on the choice of the vector potential $-\vec{\theta}$. The vector potential, in fact, is not unique and can be changed without changing the curl. The states and the flow $\vec{S}$, instead, are physical and are uniquely defined by the evolution at hand.

The other interesting part is that the Hamiltonian is the time component of the vector potential. Therefore, under a coordinate transformation, it must change like the time component of a covector. This is one of the features of special relativity and we have already found it without practically doing any work.

The last bit that I want to highlight is that the minimization of the path integral in terms of the Lagrangian is only possible if the Lagrangian can be expressed in terms of position and velocity. But the minimization in terms of the line integral of the vector potential is \textbf{always} possible. So the insight that we gathered is more general and it shows how classical mechanics is better understood in the extended phase space.

\end{document}

### 2017-05-26 Blog/04-LagrangianSubsetOfHamiltonian.tex
Final review before posting

@@
 TL;DR Lagrangians must be convex while Hamiltonians may not.
-Lagrangian systems are Hamiltonian systems where the momentum is a function $p=p(x,v)$ of position and velocity, strictly increasing with respect to velocity.
+Lagrangian systems are Hamiltonian systems where the momentum is a function $p=p(x,v)$ strictly increasing with respect to velocity (i.e.
+an increase in momentum yields an increase in velocity).
 
@@
 
-We'll see that all the Lagrangians that lead to solvable equations of motion admit a Hamiltonian (i.e.
+What happens is that that all the Lagrangians that lead to solvable equations of motion admit a Hamiltonian (i.e.
 they must be convex/hyper-regular) but not all Hamiltonians that lead to solvable equations of motions admit a Lagrangian (i.e.
@@
 But the right side are just the derivatives of $H$, which are differentiable because we assumed $H$ is twice differentiable.
-All differentiable functions are Lipschitz continuous therefore if $H$ is twice differentiable the initial value problem is always defined.
+All differentiable functions are Lipschitz continuous therefore if $H$ is twice differentiable the equations have a unique solution.
 
@@
 $H=0$): we simply get $x(t)=x_0$ and $p(t)=p_0$.
-Contrast this with setting the Lagrangian to zero.
+The Legendre transformation would give a zero Lagrangian (i.e.
+$L=0$).
 The action for all paths would be equal to $0$, all paths minimize it: there is no unique solution.
@@
 An intuitive way to remember this: Lagrangian systems are those Hamiltonian system for which each state can be recognized just by looking at its trajectory.
-If two states have the same trajectory (e.g.
+If two different states have the same trajectory (e.g.
 the velocity of a photon is always the same no matter the magnitude of momentum) then the system is not Lagrangian.

### 2017-05-26 Blog/05-ClassicalMechanicsInOnePost.tex
Reviewed with Christine

@@
 
-TL;DR Classical mechanics describes a divergence free flow of states.
+TL;DR Classical mechanics describes a divergence-free flow of states.
 The Hamiltonian is the time component of its vector potential.
@@
 For the longest time I didn't know what the Lagrangian was and why its integral is minimized along trajectories.
-A couple of months ago I went through TODO and found some insights in differential geometry that got me going and finally worked out a satisfactory answer.
+A couple of months ago I went through Mathematical Methods of Classical Mechanics by V.I.
+Arnold and found some insights in differential geometry that got me going and finally worked out a satisfactory answer.
 I unpacked that understanding in a way that only needs differential calculus, so that it's more understandable and gives direct physics intuition.
-If you have studied magnetostatic (i.e.
+If you have studied magnetostatics (i.e.
 the magnetic field and its vector potential) you should be able to fully follow this.
@@
 In this sense, I call this deterministic and reversible flow.
-Well, there are a dozen of other reasons to call it that, but we will see them some other time.
+Well, there are a dozen other reasons to call it that, but we will see them some other time.
 
@@
 \begin{aligned}
-\vec{\theta} &= p e_x + 0 e_p - H(x,p,t) e_t \\ 
+\vec{\theta} &= p \, e_x + 0 \, e_p - H(x,p,t) \, e_t \\ 
 &= \theta^x e_x + \theta^p e_p + \theta^t e_t 
@@
 \end{equation}
-This is actually the opposite of vector potential: we need an annoying minus sign to match with existing conventions. We have
+This is actually the opposite of the vector potential: we need an annoying minus sign to match with existing conventions. We have
 \begin{equation}
 \begin{aligned}
-\vec{S} &= \mathrm{curl}(-\theta) = - \nabla \times \vec{\theta} \\
+\vec{S} &= \mathrm{curl}(-\vec{\theta}) = - \nabla \times \vec{\theta} \\
 S^x &= \dfrac{dx}{dt} = - \partial_p \theta^t + \partial_t \theta^p = \partial_p H + \partial_t 0 = \partial_p H \\
-S^p &= \dfrac{dp}{dt} = - \partial_t \theta^q + \partial_q \theta^t = \partial_t p - \partial_q H = - \partial_q H \\
-S^t &= \dfrac{dt}{dt} = - \partial_q \theta^p + \partial_p \theta^q = - \partial_q 0 + \partial_p p = 1
+S^p &= \dfrac{dp}{dt} = - \partial_t \theta^x + \partial_x \theta^t = -\partial_t p - \partial_x H = - \partial_x H \\
+S^t &= \dfrac{dt}{dt} = - \partial_x \theta^p + \partial_p \theta^x = - \partial_x 0 + \partial_p p = 1
 \end{aligned}
@@
 
-The line integral of a vector field $\vec{\theta}$ has an interesting, yet little known property: it is stationary stationary along the field lines $\gamma$ of the curl.
+The line integral of a vector field $\vec{\theta}$ has an interesting, yet little known property: it is stationary along the field lines $\gamma$ of the curl.
 That is: if we make a small change in the line $\gamma$, the line integral $\int \vec{\theta} \cdot d\vec{\gamma}$ does not change.
@@
 
-\node at (-1.5, 1.25) {$\vec{S} = -\,$curl $\vec{\theta}$};
+\node at (-1.5, 1.25) {$\nabla \times \vec{\theta}$};
 
@@
 A way to intuitively understand this is the following.
-Take a field line $\gamma$, a line that is always tangent to the vector field $\vec{S}$.
-Between two point, make a small change to it $\gamma'$.
+Take a field line $\gamma$ of the curl, a line that is tangent $\nabla \times \vec{\theta}$ at every point.
+Between two points, make a small change to it $\gamma'$.
 Take $\Sigma$ to be a surface enclosed by these lines.
-The line integral of $-\vec{\theta}$ over its contour is given by the flow of its curl $\vec{S}$ through $\Sigma$.
-If the variation is small enough, the surface $\Sigma$ is tangent to the vector field $\vec{S}$ since $\gamma$ is always tangent to it.
-The line integral of $-\vec{\theta}$ over its contour is zero.
-This means that the line integral of $-\vec{\theta}$ along $\gamma$ is equal to the one along $\gamma'$: small variations of $\gamma$ do not change the line integral.
+The line integral of $\vec{\theta}$ over its contour $\partial \Sigma$ is given by the flow of its curl through $\Sigma$ by Stokes theorem.
+
+\begin{equation}
+\int_{\partial \Sigma} \vec{\theta} \cdot d\vec{\gamma} = \int_{\Sigma} \nabla \times \vec{\theta}  \cdot d\vec{\sigma}
+\end{equation}
+
+But remember: $\gamma$ is always tangent to $\nabla \times \vec{\theta}$ by construction.
+If the variation is small enough, the surface $\Sigma$ is tangent to $\nabla \times \vec{\theta}$ as well.
+The line integral of $\vec{\theta}$ over the contour $\partial \Sigma$ is zero.
+This means that the line integral of $\vec{\theta}$ along $\gamma$ is equal to the one along $\gamma'$: small variations of $\gamma$ do not change the line integral.
 
@@
 
-The last bit that I want to highlight is that the minimization of the path integral in terms of the Lagrangian is only possible if the Lagrangian can be expressed in terms of position and velocity.
-But the minimization in terms of the line integral of the vector potential is \textbf{always} possible.
+The last bit that I want to highlight is that the minimization of the path integral in terms of the Lagrangian is only possible if $p \frac{dx}{dt} - H$ can be expressed in terms of position and velocity.
+But the minimization in terms of position and momentum is \textbf{always} possible.
 So the insight that we gathered is more general and it shows how classical mechanics is better understood in the extended phase space.

### 2017-06-04 Blog/06-WhatIsMass.tex
First pass at what is mass post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{What is mass?}

TL;DR .

In this post I want to look a bit more closely at the concept of mass. There are two sides to it: inertial mass (related to $F=ma$, the resistance of the body to acceleration) and gravitational mass (related to how much an object pushes or is pushed by other objects through gravity). I will only talk about the first type as I still don't quite understand what the second is and why should it be the same thing as the first. Sometimes I feel like I am staring right at the answer but I still can't see. Anyways...

If you open a physics textbook, inertial mass is defined as the resistance of an object to being accelerated. That is, $F=ma$: if I apply the same force, the more massive object is harder to accelerate. Now, that does not quite work.

According to this notion the lesser the mass, the easier is to change the velocity of a body. Now consider the photon: it has no mass, so it should be very very easy to change it's velocity. An object of infinite mass, well, that's infinitely hard... but one that is without mass, piece of cake! Alas, the photon velocity is $c$ and cannot be changed: you cannot accelerate (or decelerate) a photon. So, that understanding of inertial mass does not quite hold.

Is there a better one? I think so and here is the gist. A range of velocity $\Delta v$ also represents a range of possible states, all the ones corresponding to the different initial conditions for velocity. But, the count of these possible states is not the same: it is proportional to the mass. That is $\dfrac{\Delta p}{\Delta v} = m$ tells us how many possible states correspond to the range of velocity. A more massive body, then, is harder to accelerate because it has to go through more states to achieve the same change in velocity. A less massive body, on the other hand, is easier to accelerate because it has fewer states to go through. A massless body, instead, cannot change velocity: no change in state $\Delta p$ will ever correspond to a change $\Delta v$: an interval of velocity is not a valid range of initial conditions for a massless particle.

It's a bit of a different perspective, that I think maps better to both the math and the physics, but let's go through some details to better understand how it works.

\section{Getting the math right}

The first order of business if to find a mathematical definition that works in all cases. $F=ma$ does not because massless particles do not accelerate: they travel at the same constant speed. $p=mv$ also does not work: the momentum of a massless particle is not zero and for a charged particle $p=mv + qA$ where $q$ is the charge and $A$ the vector potential.

We can, instead, use the following:
\begin{equation}
\label{simpleMass}
\frac{\partial p}{\partial v} = m
\end{equation}
This works for a massive charged particle, $\frac{\partial}{\partial v} (mv + qA) = m$, and for a the photon, $\frac{\partial}{\partial v} (\hbar k) = 0$ since the momentum does not depend on the velocity. In the four-dimensional case we have:
\begin{equation}
\label{advancedMass}
\frac{\partial p_\alpha}{\partial u^\beta} = m \, g_{\alpha \beta}
\end{equation}
which is interesting because it already tie in the metric tensor, i.e. the geometry of space time, in the expression. So, maybe the link between inertial and gravitational mass is already staring at me in the face... but I still don't get it.

At any rate, mathematically we can define mass as the constant that relates changes in velocity to changes in momentum. Now, what does that mean physically?

\section{Counting possibilities}

Let's go a few step backwards and ask some seemingly unrelated questions. Suppose I have a ball and five empty boxes. I want to put the ball in one of the boxes. How many possibilities do I have? The answer is simple: five.

If I get any set of discrete choices, the number of possibilities is always equal to the number of discrete choices. But what happens in the continuous case? Suppose I have a ball and a range of $1$ m. How many possibilities are there? The obvious answer would be: infinite. There are two problems with the answer: it leads to physical absurds and it is not mathematically consistent with other definitions.

Physically, suppose I increase the range to $1$ Km: that is also an infinite many possibilities. So, I have as many possibilities in $1$ m than $1$ Km. That does not sound right: I should have $1000$ more. Mathematically, the topology that we ascribe to continuous space (i.e. manifolds) is the open set topology: this means that points are not topologically distinguishable. If we can't distinguish points, we also cannot count them. (More technically, the choice of the topology already sets the measure through the Riesz representation theorem.)

So, how many possibilities are there in a range of $1$ m? $1$ m worth. How many possibilities are there in a range of $1$ Km? 1000 times more than in $1$ m. This gives an interesting insight: all numeric measurement can be seen as counting possibilities: $1$ Kelvin defines a number of possible temperatures, $1$ Pascal a number of possible pressures and so on. This relates to the importance of linearity in the definition of our unit system and therefore in all of physics: combining two intervals of $1$ meters should give us double the possibilities.

In general, suppose that I have one quantity $\hat{x}=\hat{x}(x)$ that is a linear function of another, how does the count of possibility transform? It is:
\begin{equation}
\label{countTransform}
\Delta \hat{x} = \frac{\partial \hat{x}}{\partial x} \Delta x
\end{equation}
This also gives us another insight. The partial derivative can be seen as the coefficient that transforms the count of possibilities from one unit (or quantity) to another. Relationships between different variables, then, must be differentiable if we want to map possibilities between them.

This is all very interesting: but does it have to do with mass?

\section{Counting states and trajectories}

For particle mechanics, we have at least two things we may want to count: states and trajectories. States are identified by position and momentum (i.e. the state variables) while trajectories are identified by position and velocity (i.e. the initial conditions). Since position is the same for both, the relationship between momentum and velocity is enough to tell us how the count changes. That is, if we combine \eqref{simpleMass} and \eqref{countTransform} (and skip some mathematical details) we have:
\begin{equation}
\label{stateCountTransform}
\Delta p = \frac{\partial p}{\partial v} \Delta v = m \Delta v
\end{equation}

In other words: inertial mass transforms the count of possible trajectories into the count of possible states. A more massive body has fewer trajectories within the same possible states. This makes intuitive sense: a force changes the state of a system, goes through the same possible states, but accelerates different bodies differently depending on the ratio between states and trajectories. This reasoning extends to massless particles, like the photon. If the particle has no mass, momentum and velocity are independent variables. Changes is momentum will not produce any change in velocity.

The key here is that the derivative of the inverse is not the derivative of the inverse. In fact if $p$ is not a function of $v$ then $v$ is not a function of $p$: $\dfrac{\partial p}{\partial v} = \dfrac{\partial v}{\partial p} = 0$. In the introduction, in fact, we observed how the photon is impossible to accelerate, which should suggest the mass is infinite. Well, in a way, the mass and its inverse are both zero. It looks like that because what we are really saying is that the magnitude of momentum and velocity are independent variables: looking at the trajectory of a photon tells us nothing about the magnitude of its momentum.

\section{Conclusion}

We have seen that the inertial mass can be given a more general physical meaning as the constant that relates the count of possible states to the count of possible trajectories. This gives a concept that works in a more general context.

Not everything, though, is nice and neat. In the massless case the velocity and the momentum are not completely independent: the direction is the same. How does that fit into the picture? In \eqref{advancedMass} the metric tensor does come into place: how does the space-time curvature (i.e. gravity) affect the relationship? The same relationship implies that momentum is a linear function of velocity: does it have to be so and why?

It would make sense if, in some way, the geometry of space-time affects the possibility count of trajectories: gravity would be the "apparent" force that results from such a correction. But that's just speculation: I don't see the math yet. I do hope one day to find answers to these questions and that they will give more insight on why inertial mass and gravitational forces are related. Unfortunately, I can never predict when...

\end{document}

### 2017-06-14 Blog/05-ClassicalMechanicsInOnePost.tex
Minor improvements

@@
 A couple of months ago I went through Mathematical Methods of Classical Mechanics by V.I.
-Arnold and found some insights in differential geometry that got me going and finally worked out a satisfactory answer.
+Arnold and found some insights into differential geometry that got me going and finally worked out a satisfactory answer.
 I unpacked that understanding in a way that only needs differential calculus, so that it's more understandable and gives direct physics intuition.
@@
 
-If you remember your vector calculus, you know that we can define a vector potential for a divergence free field. We therefore define
+If you remember your vector calculus, you know that we can define a vector potential for a divergence-free field. We therefore define
 \begin{equation}
@@
 A way to intuitively understand this is the following.
-Take a field line $\gamma$ of the curl, a line that is tangent $\nabla \times \vec{\theta}$ at every point.
+Take a field line $\gamma$ of the curl, a line that is tangent to $\nabla \times \vec{\theta}$ at every point.
 Between two points, make a small change to it $\gamma'$.
 Take $\Sigma$ to be a surface enclosed by these lines.
-The line integral of $\vec{\theta}$ over its contour $\partial \Sigma$ is given by the flow of its curl through $\Sigma$ by Stokes theorem.
+The line integral of $\vec{\theta}$ over its contour $\partial \Sigma$ is given by the flow of its curl through $\Sigma$ by the Stokes theorem.
 

### 2017-06-14 Blog/06-WhatIsMass.tex
Minor improvements

@@
 In this post I want to look a bit more closely at the concept of mass.
-There are two sides to it: inertial mass (related to $F=ma$, the resistance of the body to acceleration) and gravitational mass (related to how much an object pushes or is pushed by other objects through gravity).
-I will only talk about the first type as I still don't quite understand what the second is and why should it be the same thing as the first.
-Sometimes I feel like I am staring right at the answer but I still can't see.
-Anyways...
+We are mainly going to talk about inertial mass (the one used in $F=ma$) as I still don't quite understand gravitational mass (the one related to gravitational forces and space-time curvature).
 
@@
 That is, $F=ma$: if I apply the same force, the more massive object is harder to accelerate.
-Now, that does not quite work.
+Now, that does not quite work in general.
 
 According to this notion the lesser the mass, the easier is to change the velocity of a body.
-Now consider the photon: it has no mass, so it should be very very easy to change it's velocity.
-An object of infinite mass, well, that's infinitely hard...
-but one that is without mass, piece of cake!
-Alas, the photon velocity is $c$ and cannot be changed: you cannot accelerate (or decelerate) a photon.
+Now consider the photon: its speed is the constant $c$ and cannot be changed.
+In other words: the photon is infinitely hard to accelerate.
+It should have infinite mass, right?
+Actually no: it has zero mass.
 So, that understanding of inertial mass does not quite hold.
@@
 I think so and here is the gist.
-A range of velocity $\Delta v$ also represents a range of possible states, all the ones corresponding to the different initial conditions for velocity.
-But, the count of these possible states is not the same: it is proportional to the mass.
-That is $\dfrac{\Delta p}{\Delta v} = m$ tells us how many possible states correspond to the range of velocity.
+A small range of velocity $dv$ defines a set of possible trajectories for a particle: the ones corresponding to the different initial conditions for velocity.
+This will correspond to a set of possible states defined by a range of momentum $dp$.
+The ratio between the count of possible states and the count of possible trajectories, $\dfrac{\partial p}{\partial v} = m$, tells us how many possible states are there in each unit of velocity.
 A more massive body, then, is harder to accelerate because it has to go through more states to achieve the same change in velocity.
 A less massive body, on the other hand, is easier to accelerate because it has fewer states to go through.
-A massless body, instead, cannot change velocity: no change in state $\Delta p$ will ever correspond to a change $\Delta v$: an interval of velocity is not a valid range of initial conditions for a massless particle.
+For a massless body, instead, a range of velocity will not correspond to a range of possible states.
+$p$ is not a function of $v$ and therefore a change in state (i.e.
+momentum) will not correspond to a change in trajectory (i.e.
+velocity).
 
@@
 This works for a massive charged particle, $\frac{\partial}{\partial v} (mv + qA) = m$, and for a the photon, $\frac{\partial}{\partial v} (\hbar k) = 0$ since the momentum does not depend on the velocity.
-In the four-dimensional case we have:
+For those familiar with relativity, in the four-dimensional case we have:
 \begin{equation}
@@
 \end{equation}
-which is interesting because it already tie in the metric tensor, i.e.
+which is interesting because it already ties in the metric tensor, i.e.
 the geometry of space time, in the expression.
@@
 
-This is all very interesting: but does it have to do with mass?
+This is all very interesting: but what does it have to do with mass?
 

### 2017-06-22 Blog/06-WhatIsMass.tex
Reviewed with Christine

@@
 
-TL;DR .
+TL;DR The inertial mass tells us how many states are there within a given range of trajectories.
+More massive bodies are harder to accelerate because they have to go through more states to reach the same change in trajectory.
 
@@
 
-It's a bit of a different perspective, that I think maps better to both the math and the physics, but let's go through some details to better understand how it works.
+It's a bit of a different perspective, which I think maps better to both the math and the physics, but let's go through some details to better understand how it works.
 
@@
 $F=ma$ does not because massless particles do not accelerate: they travel at the same constant speed.
-$p=mv$ also does not work: the momentum of a massless particle is not zero and for a charged particle $p=mv + qA$ where $q$ is the charge and $A$ the vector potential.
+$p=mv$ does not work either: the momentum of a massless particle is not zero and for a charged particle $p=mv + qA$ where $q$ is the charge and $A$ the vector potential.
 
@@
 \end{equation}
-This works for a massive charged particle, $\frac{\partial}{\partial v} (mv + qA) = m$, and for a the photon, $\frac{\partial}{\partial v} (\hbar k) = 0$ since the momentum does not depend on the velocity.
+This works for a massive charged particle, $\frac{\partial}{\partial v} (mv + qA) = m$, and for a photon, $\frac{\partial}{\partial v} (\hbar k) = 0$ since the momentum does not depend on the velocity.
 For those familiar with relativity, in the four-dimensional case we have:
@@
 which is interesting because it already ties in the metric tensor, i.e.
-the geometry of space time, in the expression.
-So, maybe the link between inertial and gravitational mass is already staring at me in the face...
+the geometry of space-time, in the expression.
+So, maybe the link between inertial and gravitational mass is already staring me in the face...
 but I still don't get it.
@@
 
-Let's go a few step backwards and ask some seemingly unrelated questions.
+Let's go a few steps backwards and ask some seemingly unrelated questions.
 Suppose I have a ball and five empty boxes.
@@
 The obvious answer would be: infinite.
-There are two problems with the answer: it leads to physical absurds and it is not mathematically consistent with other definitions.
+There are two problems with the answer: it leads to physical absurdities and it is not mathematically consistent with other definitions.
 
-Physically, suppose I increase the range to $1$ Km: that is also an infinite many possibilities.
-So, I have as many possibilities in $1$ m than $1$ Km.
-That does not sound right: I should have $1000$ more.
-Mathematically, the topology that we ascribe to continuous space (i.e.
-manifolds) is the open set topology: this means that points are not topologically distinguishable.
-If we can't distinguish points, we also cannot count them.
-(More technically, the choice of the topology already sets the measure through the Riesz representation theorem.)
+Physically, suppose I increase the range to $1$ Km: that is also infinitely many possibilities.
+So, I have as many possibilities in $1$ m as $1$ Km.
+That does not sound right: I should have $1000$ times more.
+If you happen to have studied topology, the topology that we ascribe to continuous space (i.e.~manifolds) is the open set topology.
+This means that isolated points are of measure zero (not one) and finite ranges have finite measure (not infinite).
+That is: the choice of the topology already sets the measure through the Riesz representation theorem.
 
@@
 1000 times more than in $1$ m.
-This gives an interesting insight: all numeric measurement can be seen as counting possibilities: $1$ Kelvin defines a number of possible temperatures, $1$ Pascal a number of possible pressures and so on.
-This relates to the importance of linearity in the definition of our unit system and therefore in all of physics: combining two intervals of $1$ meters should give us double the possibilities.
+This gives an interesting insight: all numeric measurement can be seen as counting possibilities: a range of $1$ Kelvin defines a number of possible temperatures, a range of $1$ Pascal a number of possible pressures and so on.
+This relates to the importance of linearity in the definition of our unit system and therefore in all of physics: combining two intervals of $1$ meter should give us double the possibilities.
 
-In general, suppose that I have one quantity $\hat{x}=\hat{x}(x)$ that is a linear function of another, how does the count of possibility transform? It is:
+In general, suppose that I have one quantity $\hat{x}=\hat{x}(x)$ that is a linear function of another, how does the count of possibilities transform? It is:
 \begin{equation}
@@
 For particle mechanics, we have at least two things we may want to count: states and trajectories.
-States are identified by position and momentum (i.e.
-the state variables) while trajectories are identified by position and velocity (i.e.
-the initial conditions).
+States are identified by position and momentum (i.e.~the state variables) while trajectories are identified by position and velocity (i.e.~the initial conditions).
 Since position is the same for both, the relationship between momentum and velocity is enough to tell us how the count changes.
@@
 If the particle has no mass, momentum and velocity are independent variables.
-Changes is momentum will not produce any change in velocity.
+Changes in momentum will not produce any change in velocity.
 
-The key here is that the derivative of the inverse is not the derivative of the inverse.
+The key here is that the derivative of the inverse is not the inverse of the derivative.
 In fact if $p$ is not a function of $v$ then $v$ is not a function of $p$: $\dfrac{\partial p}{\partial v} = \dfrac{\partial v}{\partial p} = 0$.
@@
 How does that fit into the picture?
-In \eqref{advancedMass} the metric tensor does come into place: how does the space-time curvature (i.e.
-gravity) affect the relationship?
+In \eqref{advancedMass} the metric tensor does come into place: how does the space-time curvature (i.e.~gravity) affect the relationship?
 The same relationship implies that momentum is a linear function of velocity: does it have to be so and why?
@@
 It would make sense if, in some way, the geometry of space-time affects the possibility count of trajectories: gravity would be the "apparent" force that results from such a correction.
-But that's just speculation: I don't see the math yet.
-I do hope one day to find answers to these questions and that they will give more insight on why inertial mass and gravitational forces are related.
+I do hope one day to find answers to these questions and that they will give more insight into why inertial mass and gravitational forces are related.
 Unfortunately, I can never predict when...

### 2017-07-04 Blog/06-WhatIsMass.tex
Better summary

@@
 
-TL;DR The inertial mass tells us how many states are there within a given range of trajectories.
-More massive bodies are harder to accelerate because they have to go through more states to reach the same change in trajectory.
+TL;DR Inertial mass tells us how many states are there within a given range of velocity.
+More massive bodies are harder to accelerate because they have to go through more states to reach the same change in velocity.
 

### 2017-07-05 Articles/TopologyAndExperimentalDistinguishability/article.tex
Add scheleton for article

\documentclass[review]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Topology and its Applications}
\bibliographystyle{elsarticle-num}

\begin{document}

\begin{frontmatter}

\title{Topology and physical distinguishability}

%% Group authors per affiliation:
\author{Gabriele Carcassi\fnref{myfootnote}}
\address{Somewhere, Ann Arbor}
\fntext[myfootnote]{Footnote?}

%% or include affiliations in footnotes:
\author[mymainaddress,mysecondaryaddress]{University of Michigan}
\ead[url]{www.umich.edu}

\author[mysecondaryaddress]{Secondary\corref{mycorrespondingauthor}}
\cortext[mycorrespondingauthor]{Corresponding author}
\ead{carcassi@umich.edu}

\address[mymainaddress]{Somewhere else, Ann Arbor}
\address[mysecondaryaddress]{Somewhere instead, Ann Arbor}

\begin{abstract}
TODO
\end{abstract}

\begin{keyword}
topological spaces \sep other keyword
\MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

\section{Experimental observations}

\section{Experimental distinguishability}

\section{Experimental relationships}

\section{Conclusion}

\bibliography{bibliography}

\end{document}

### 2017-07-05 Articles/TopologyAndExperimentalDistinguishability/article.tex
added relevant book material

@@
 \usepackage{lineno,hyperref}
+
+\usepackage{amsmath, amsthm, amsfonts}
+\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}
+
+\theoremstyle{plain}% default 
+\newtheorem{thm}{Theorem}[section] 
+\newtheorem{lem}[thm]{Lemma} 
+\newtheorem{prop}[thm]{Proposition} 
+\newtheorem{cor}{Corollary}
+
+ \theoremstyle{definition}
+ \newtheorem{defn}{Definition}[section]
+\newtheorem{exmp}{Example}[section]
+\theoremstyle{remark}
+\newtheorem*{rem}{Remark}
 \modulolinenumbers[5]
@@
 
+\begin{defn}
+	A \textbf{statement} $\mathsf{s}$ is a declarative sentence that is either true or false, as in classical logic.
+\end{defn}
+
+\begin{defn}
+	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure that can terminate successfully in a finite amount of time and that can be stopped at any time.
+\end{defn}
+
+TODO: in principle, time can be defined as a monoid. This would prevent circular arguments to some extent.
+
+\begin{defn}
+	An \textbf{experimental observation} is a tuple $\mathsf{o} = \llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is always successful.
+The experimental observation is \textbf{verified} if the statement is true.
+\end{defn}
+
+\begin{defn}
+	An \textbf{experimental counter-observation} is a tuple $\mathsf{o}^c= \llbracket \mathsf{s}, \mathsf{e} \rrbracket$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is false if and only if the experimental test is always successful.
+The experimental counter-observation is \textbf{refuted} if the statement is false.
+\end{defn}
+
+We also define some basic relationships between experimental observations.
+
+\begin{defn}
+	Given two experimental observations $\mathsf{o}_1$ and $\mathsf{o}_2$:
+	\begin{itemize}
+	\item \textbf{implication}: $\mathsf{o}_1 \implies \mathsf{o}_2$ if $\mathsf{o}_2$ is verified whenever $\mathsf{o}_1$ is verified
+	\item \textbf{equivalence}: $\mathsf{o}_1 = \mathsf{o}_2$ if $\mathsf{o}_1 \implies \mathsf{o}_2$ and $\mathsf{o}_2 \implies \mathsf{o}_1$
+	\item \textbf{mutual exclusion}: $\mathsf{o}_1 \asymp \mathsf{o}_2$ ($\mathsf{o}_1$ excludes $\mathsf{o}_2$) if one is not verified whenever the other one is
+	\end{itemize}
+\end{defn}
+TODO: add a symbol for ``not mutually exclusive". I couldn't find the $\asymp$ with a slash through it.
+
+A note on the language: even if $\mathsf{o}_1$ implies $\mathsf{o}_2$ this does not give us a way to verify $\mathsf{o}_2$.
+There may be cases where the $\mathsf{o}_2$ is verified and $\mathsf{o}_1$ is not.
+Therefore it is improper to say that one observation verifies another: only experimental tests verify observations.
+
+We define the following logical operations on experimental observations.
+Taken as whole, they define the \textbf{algebra of experimental observations}.
+With each definition, we will prove that the resulting object is an experimental (counter-)observation.
+
+\begin{defn}
+	The \textbf{negation or logical NOT} of an experimental observation $\mathsf{o}=\llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ is the experimental counter-observation $\neg \mathsf{o}=\llbracket\neg \mathsf{s}, \mathsf{e}\rrbracket$ where the negation of the statement $\mathsf{s}$ can be refuted by the experimental test $\mathsf{e}$.
+\end{defn}
+
+\begin{proof}
+	We show that the negation exists and is well defined.
+Let $\mathsf{o}=\llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ be an experimental observation.
+The experimental test $\mathsf{e}$ succeeds if and only if $\mathsf{s}$ is true or, equivalently, if and only if $\neg \mathsf{s}$ is false.
+So the tuple $\llbracket\neg \mathsf{s}, \mathsf{e}\rrbracket$ forms an experimental counter-observation.
+\end{proof}
+
+\begin{defn}
+	The \textbf{conjunction or logical AND} of a finite collection  of experimental observations $\{\mathsf{o}_i\}_{i=1}^{n}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{n}$ is the experimental observation $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigwedge\limits_{i=1}^{n} \mathsf{s}_i$ is the conjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ is the experimental test that successfully terminates if and only if all $\{\mathsf{e}_i\}_{i=1}^{n}$ successfully terminate.
+\end{defn}
+
+\begin{proof}
+	We show that the conjunction exists and is well defined. Let $\mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ be the experimental procedure defined as follows:
+	\begin{enumerate}
+	\item for each $i=1..n$ run the test $\mathsf{e}_i$
+	\item if all tests terminated successfully terminate successfully.
+	\end{enumerate}
+	The experimental procedure so defined terminates successfully if and only if all $\mathsf{e}_i$ terminate successfully.
+It will do so in finite time as each $\mathsf{e}_i$ succeeds in finite time and there only a finite number of them.
+Therefore $\mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ is an experimental test that is successful if and only if all statements $\{\mathsf{s}_i\}_{i=1}^{n}$ are successful.
+So $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis\bigwedge\limits_{i=1}^{n} \mathsf{s}_i, \mathsf{e}_{\wedge}(\mathsf{e}_i)\rrparenthesis$ is an experimental observation.
+\end{proof}
+
+\begin{defn}
+	The \textbf{disjunction or logical OR} of a countable (finite or infinite) collection of experimental observations $\{\mathsf{o}_i\}_{i=1}^{\infty}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{\infty}$ is the experimental observation $\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigvee\limits_{i=1}^{\infty} \mathsf{s}_i$ is the disjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ is the experimental test that successfully terminates if and only if at least one experimental test in $\{\mathsf{e}_i\}_{i=1}^{\infty}$ successfully terminates.
+\end{defn}
+
+\begin{proof}
+	We show that the disjunction exists and is well defined. Let $\mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ be the experimental procedure defined as follows:
+	\begin{enumerate}
+	\item initialize $n$ to 1
+	\item for each $i=1..n$
+	\begin{enumerate}
+		\item run the test $\mathsf{e}_i$ for $n$ seconds
+		\item if $\mathsf{e}_i$ terminated successfully, terminate successfully
+	\end{enumerate}
+	\item increment $n$ and go to step 2
+	\end{enumerate}
+	Suppose there exists an $i \in \mathbb{Z}^+$ such that $\mathsf{e}_i$ will terminate successfully, then the above procedure will eventually run that test for a time long enough and terminate successfully.
+It will do so in finite time as it will have run finitely many tests finitely many times each for a finite amount of time.
+Therefore $\mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ is an experimental test that is successful if and only if at least one statement in $\{\mathsf{s}_i\}_{i=1}^{n}$ is successful.
+So $\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i =\llparenthesis\bigvee\limits_{i=1}^{\infty} \mathsf{s}_i, \mathsf{e}_{\wedge}(\{\mathsf{e}_i\}_{i=1}^{\infty})\rrparenthesis$ is an experimental observation.
+\end{proof}
+
+For technical reasons, we will need the following:
+
+\begin{defn}
+The \textbf{empty observation} is any experimental observation of the form $\llparenthesis \mathsf{s}\wedge \neg\mathsf{s},\mathsf{e}\rrparenthesis$. Note that this will always fail. 
+\end{defn}
+
+\section{Experimental domain}
+
+\begin{defn}
+	An \textbf{generalized experimental domain} is a set of observations closed under finite conjunction and countable disjunction.
+Any subset that can generate all others via finite conjunction and countable disjunction is called a \textbf{sub-basis} of the domain.
+A \textbf{basis} is a subset of the observations which can generate all others by countable disjunctions.
+\end{defn}
+
+Clearly, given a sub-basis one can generate a basis by taking all finite conjunctions of the observations.
+Note that any infinite sub-basis will generate a basis of the same cardinality.
+We claim that it is only possible to verify all members of an experimental domain with a countable (sub-)basis even given infinite time.
+
+\begin{prop}
+Let $\mathcal{S}$ be an experimental domain.
+Then it is possible to verify all experimental observations $\mathsf{o}\in S$ in infinite time if and only if there exists a countable basis (equivalently, sub-basis) $\mathcal{B}$ of $\mathcal{S}$.
+\end{prop}
+\begin{proof}
+If there exists a countable basis $\mathcal{B}$, then given infinite time one can test all observations in $\mathcal{B}$.
+Given all truth values for elements of the basis, one can deduce the truth values for the other observations in $\mathcal{S}$ (again using infinite time) by computing the appropriate conjunctions and disjunctions.
+
+If there does not exist a countable basis, then by definition there does not exist a sequence of experimental observations in $\mathcal{S}$ from which one can deduce all other observations in $\mathcal{S}$.
+Hence it is impossible to verify all members of $\mathcal{S}$.
+\end{proof}
+
+In light of the above result, we have the following definition, on which the remainder of this chapter will focus.
+
+\begin{defn}
+An \textbf{experimental domain} is a set of observations closed under finite conjunction and countable disjunction which has a countable basis and contains the empty observation. 
+\end{defn}
+
 \section{Experimental distinguishability}
+
+We will now use our notions of experimental observations and experimental domains to construct a new space with a topology which will more rigorously capture our notion of ``exact" possibilities.
+We will begin with a given experimental domain $\mathcal{S}$.
+
+\begin{defn}
+A \textbf{generalized exact outcome} for $\mathcal{S}$ is a sequence of experimental observations $\mathcal{P} = \{\mathsf{o}_i\}_{i=1}^{\infty}\subset\mathcal{S}$ with the following property.
+For all experimental observations $\mathsf{o}\in\mathcal{S}$, there exists some positive integer $N$ such that for all integers $n>N$, one of the following must happen:
+\begin{enumerate}
+\item $\mathsf{o}_n \asymp \mathsf{o}$ 
+\item $\mathsf{o}_n \implies \mathsf{o}$
+\end{enumerate}
+\end{defn}
+
+\begin{defn}
+An \textbf{exact outcome} for $\mathcal{S}$ is a generalized exact outcome $\mathcal{P} = \{\mathsf{o}_i\}_{i=1}^{\infty}$ such that for all $i\geq1$, we have $\mathsf{o}_{i+1}\implies\mathsf{o}_i$.
+\end{defn}
+Note that given a generalized exact outcome, one can build an exact outcome by simply replacing each $\mathsf{o}_n$ with the (finite) conjunction $\bigvee_{i=1}^n\mathsf{o}_i$.
+We will often denote the collection of (generalized) exact outcomes by $\mathcal{C}$.
+Next, we note that different sequences can result in ``equivalent" outcomes.
+
+\begin{defn}
+We say two exact outcomes $\mathcal{P} = \{\mathsf{o}_i\}_{i=1}^{\infty}$ and $\mathcal{P}' = \{\mathsf{o}'_i\}_{j=1}^{\infty}$ are \textbf{equivalent}, denoted $\mathcal{P}\sim\mathcal{P}'$, if for all $i,j$, we have that $\mathsf{o}_i$ and $\mathsf{o}'_j$ are not mutually exclusive.
+\end{defn}
+
+Note that given a generalized exact outcome, the exact outcome built from it will be equivalent.
+Importantly, this notion of equivalence is in fact an equivalence relation (i.e.
+a relation with reflexivity, symmetry, and transitivity):
+
+\begin{prop}
+The equivalence $\sim$ above is an equivalence relation.
+\end{prop}
+\begin{proof}
+Let $\mathcal{P},\mathcal{P}',\mathcal{P}''$ be exact outcomes. Reflexivity and symmetry are immediate since not being mutually exclusive is reflexive and symmetric.
+
+For transitivity, suppose $\mathcal{P}\sim\mathcal{P}'$ and $\mathcal{P}'\sim\mathcal{P}''$, and pick some positive integers $m$ and $n$.
+We will show that $\mathsf{o}_m$ is not mutually exclusive to $\mathsf{o}''_n$.
+By hypothesis, for every $\mathsf{o}'_j\in\mathcal{P}'$, we have $\mathsf{o}_m$ and $\mathsf{o}'_j$ are not mutually exclusive and $\mathsf{o}'_j$ and $\mathsf{o}''_n$ are not mutually exclusive.
+Because $\mathcal{P}'$ is an exact outcome and neither $\mathsf{o}_m$ nor $\mathsf{o}''_n$ are ever mutually exclusive from any $\mathsf{o}'_j$, we may select an integer $J$ large enough so that $\mathsf{o}'_J\implies\mathsf{o}_m$ and $\mathsf{o}'_J\implies\mathsf{o}''_n$.
+Hence whenever $\mathsf{o}'_J$ is verified, both $\mathsf{o}_m$ and $\mathsf{o}''_n$ are verified, so they are not mutually exclusive.
+Because this holds for all positive integers $m,n$, we have that $\mathcal{P}_1\sim\mathcal{P}_3$.
+\end{proof}
+
+Next, we will define our universe of discourse.
+
+\begin{defn}
+	A \textbf{set of possibilities} $X$ for an experimental domain $\mathcal{S}$ is the collection $\mathcal{C}$ of all exact outcomes for $\mathcal{S}$, modulo the equivalence relation $\sim$ defined above.
+That is, $X=\mathcal{C}/\sim$.
+\end{defn}
+
+Equivalently, a set of possibilities $X$ for $\mathcal{S}$ may be specified by choosing one exact outcome from each equivalence class.
+We will now provide a notion of containment for exact outcomes being ``within" an experimental observation.
+
+\begin{defn}
+An exact outcome $\mathcal{P}=\{\mathsf{o}_i\}_{i=1}^{\infty}$ for experimental domain $\mathcal{S}$ is said to be an element of an experimental observation $\mathsf{o}\in\mathcal{S}$ if there exists some $N$ such that for all $n>N$, we have $\mathsf{o}_n\implies\mathsf{o}$.
+We denote this as $\mathcal{P}\in\mathsf{o}$.
+\end{defn}
+
+In light of the above definition, we will refer to experimental observations either as observations or as sets.
+
+\begin{prop}
+Let $\mathcal{P}=\{\mathsf{o}_i\}_{i=1}^{\infty}$ be an exact outcome which is an element of observation $\mathsf{o}$.
+Suppose $\mathcal{P}'$ is another exact outcome $\mathcal{P}'=\{\mathsf{o}'_j\}_{j=1}^{\infty}$ such that $\mathcal{P}\sim\mathcal{P}'$.
+Then $\mathcal{P}'\in\mathsf{o}$.
+\end{prop}
+\begin{proof}
+Suppose not - then we must be in case (1) of the definition of (generalized) exact outcome.
+Then there exists some $N_1$ such that for all $n>N_1$, we have $\mathsf{o}'_n\asymp\mathsf{o}$.
+By hypothesis, there exists some $N_2$ such that for all $m>N_2$, we have $\mathsf{o}_m\implies\mathsf{o}$.
+Let $N=\max(N_1,N_2)$.
+But then for all $k>N$, we have that $\mathsf{o}_k$ and $\mathsf{o}'_k$ are not mutually exclusive.
+This contradicts $\mathsf{o}'_k\asymp\mathsf{o}$.
+\end{proof}
+
+In particular, we immediately extend the definition of ``exact outcome containment" for elements of $\mathcal{S}$ to containment of equivalence classes of exact outcomes in $\mathcal{C}/\sim$.
+By the above, we can use containment of an individual exact outcome interchangeably with its equivalence class.
+In practice, most of our work will use individual exact outcomes representing an equivalence class, while the technical definitions will use their equivalence classes.
+
+\begin{prop}
+Let $\{\mathsf{o}_i\}_{i=1}^{n}$ be some collection of experimental observations.
+Then the conjunction $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis\bigwedge\limits_{i=1}^{n} \mathsf{s}_i, \mathsf{e}_{\wedge}(\mathsf{e}_i)\rrparenthesis$ contains precisely the exact outcomes that are in all of the $\mathsf{o}_i$'s.
+\end{prop}
+\begin{proof}
+Let $\mathcal{P} = \{\mathsf{o}'_i\}_{i=1}^{\infty}$ be an exact outcome, and suppose $\mathcal{P}\in\mathsf{o}_i$ for $i=1,\ldots,n$.
+Then there exists $N$ large enough so that $\mathsf{o}'_m\implies\mathsf{o}_i$ for all $m>N$ and all $i=1,\ldots,n$, which is possible since $n$ is finite.
+It follows that $\mathsf{o}'_m\implies\bigwedge\limits_{i=1}^{n} \mathsf{o}_i$ for all $m>N$, so $\mathcal{P}\in\bigwedge\limits_{i=1}^{n} \mathsf{o}_i$, as desired.
+
+Next, let $\mathcal{P}\in\bigwedge\limits_{i=1}^{n} \mathsf{o}_i$. It follows that $\mathcal{P}\in\mathsf{o}_i$ for each $i$ immediately from the definitions. 
+\end{proof}
+
+\begin{prop}
+Let $\{\mathsf{o}_i\}_{i=1}^{\infty}$ be some collection of experimental observations.
+Then the disjunction $\bigvee\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis\bigvee\limits_{i=1}^{n} \mathsf{s}_i, \mathsf{e}_{\vee}(\mathsf{e}_i)\rrparenthesis$ contains precisely the exact outcomes that are in any of the $\mathsf{o}_i$'s.
+\end{prop}
+\begin{proof}
+Let $\mathcal{P} = \{\mathsf{o}'_i\}_{i=1}^{\infty}$ be an exact outcome, and suppose $\mathcal{P}\in\mathsf{o}_k$ for some $k\geq1$.
+Then there exists $N$ large enough so that $\mathsf{o}'_m\implies\mathsf{o}_k$ for all $m>N$.
+It follows that $\mathsf{o}'_m\implies\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i$ for all $m>N$, so $\mathcal{P}\in\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i$, as desired.
+
+Next, let $\mathcal{P}\in\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i$.
+Then there exists $N$ such that for all $m>N$, we have $\mathsf{o}'_m\implies \bigvee\limits_{i=1}^{\infty} \mathsf{o}_i$.
+Pick some such $m$, and so in particular, there exists $k$ such that $\mathsf{o}'_m\implies\mathsf{o}_k$.
+By definition of (non-generalized) exact outcome, this means $\mathcal{P}\in \mathsf{o}_k$, as required.
+\end{proof}
+
+TODO: check this. the definition of experimental observation and their OR and AND are a little hard to use as needed here.
+
+TODO: maybe make a new notation for observation as a set of outcomes.
+
+TODO: a lot of the indices/notation might be a little hard to read/follow.
+
+We are now ready to prove one of the main results of this chapter.
+The following theorem links the philosophical definitions at the beginning of the chapter to the mathematical structure of a topological space.
+
+TODO: maybe add a theorem environment for this one
+
+\begin{prop}
+Let $\mathcal{S}$ be an experimental domain with (by definition countable) basis $\mathcal{B}$, and let $X = \mathcal{C}/\sim$ be the set of possibilities.
+Then $(X,\mathcal{S})$ is a Hausdorff, second-countable topological space with points $X$ and open sets $\mathcal{S}$, where union and intersection are given by conjunction and disjunction of experimental observations, respectively.
+\end{prop}
+\begin{proof}
+We first will show that the collection of sets $\mathsf{o}\in\mathcal{S}$ do indeed form a topology on $X$.
+First, the experimental observation formed by the (countable) disjunction of all observations in $\mathcal{B}$ is an observation which contains all others; that is, this set contains all of $X$.
+The empty observation contains no elements of $X$, so serves as the empty set.
+
+Next, let $\mathsf{o}_1,\mathsf{o}_2\in\mathcal{S}$.
+Then by the previous proposition about conjunction, we have that the set of exact outcomes corresponding to $\mathsf{o}_1\wedge\mathsf{o}_2$ is precisely $\mathsf{o}_1\cap\mathsf{o}_2$.
+For a countable subset of $\mathcal{S}$, the same argument shows that their union (as sets of exact outcomes) matches the disjunction (as experimental observations).
+For arbitrary collections, because $\mathcal{B}$ is countable and generates all of the experimental observations, we can rewrite an arbitrary union of the corresponding subsets of exact outcomes as a countable disjunction of experimental observations, and we are reduced to the countable case.
+
+This proves that $(X,\mathcal{S})$ is a topology.
+Further, we can see that $\mathcal{B}$ is a countable basis for the topology (because conjunction and disjunction correspond to intersection and union), so it is second-countable.
+Next, we prove that $(X,\mathcal{S})$ is Hausdorff.
+
+Let $\mathcal{P} = \{\mathsf{o}_i\}_{i=1}^{\infty}$ and $\mathcal{P}' = \{\mathsf{o}'_i\}_{j=1}^{\infty}$ be two non-equivalent exact outcomes.
+Then there exist some $m,n$ such that $\mathsf{o}_m\asymp\mathsf{o}'_n$.
+But then $\mathcal{P}\in\mathsf{o}_m$ and $\mathcal{P}'\in\mathsf{o}'_n$, which are two disjoint sets in $\mathcal{S}$.
+This completes the proof.
+\end{proof}
+
+\begin{defn}
+	A \textbf{verifiable set} $U \subseteq X$ is a subset of possibilities for which there exists an experimental observation $\mathsf{o}\in\mathcal{S}$  = (\text{``The object is in } U \text{"}, $\mathsf{e}_\in(U))$ where $\mathsf{e}_\in(U)$ is an experimental test that succeeds only if the object to identify is an element of $U$.
+\end{defn}
+
+\begin{defn}
+	A \textbf{refutable set} $U \subseteq X$ is a subset of possibilities for which there exists an experimental counter-observation $\mathsf{o} = (\text{``The object is in } U \text{"}, \mathsf{e}_{\notin}(U))$ where $\mathsf{e}_{\notin}(U)$ is an experimental test that succeeds only if the object to identify is not an element of $U$.
+\end{defn}
+
+\begin{prop}
+	The complement $U^C$ of a verifiable set $U \subseteq X$ is a refutable set.
+\end{prop}
+
+\begin{prop}
+	A Hausdorff and second countable space $X$ has at most cardinality of continuum.
+\end{prop}
+
+\begin{proof}
+	We define an injective function $F:X\to2^{\mathbb{N}}$, where $2^{\mathbb{N}}$ denotes all infinite binary sequences.
+
+	Since $X$ is second-countable, we can enumerate the countable basis $\mathcal{B}$ as $\mathcal{B} = \{B_i\}_{i=1}^{\infty}$.
+Let $2^{\mathbb{N}}$ denote all infinite binary sequences.
+We define $F:X\to2^{\mathbb{N}}$ such that $F(x) = (F(x)_i)_{i=1}^{\infty}$ is the sequence where each element is given by:
+	$$
+	F(x)_i = 
+	\begin{cases}
+	1 & x\in B_i \\
+	0 & x\notin B_i
+	\end{cases}
+	$$
+	This is an injective function.
+Suppose $x_1 \neq x_2$, then since $X$ is Hausdorff there is at least one element of the basis that contains one but not the other.\footnote{In fact, $T_0$ separability would be sufficient.} Therefore $F(x_1) \neq F(x_2)$.
+As $F$ injects $X$ into $2^{\mathbb{N}}$, we have $|X| \leq |2^{\mathbb{N}}|=|\mathbb{R}|$.
+$X$ has at most cardinality of continuum.
+\end{proof}
+
+This means that, no matter what technique we use now or in the future, the collections of elements that we can properly define experimentally are at most infinite like the continuum.
+This already gives us a first simple and basic requirement a set of mathematical objects need to pass to be of scientific interest.
+
+For example, these objects have cardinality of continuum, and therefore are good candidates:
+\begin{itemize}
+	\item Euclidean space $\mathbb{R}^n$
+	\item all continuous functions from $\mathbb{R}$ to $\mathbb{R}$
+	\item all open sets in $\mathbb{R}^n$
+	\item all subsets of $\mathbb{N}$
+\end{itemize}
+
+These, instead, have cardinality greater then continuum, and therefore are not good candidates:
+\begin{itemize}
+	\item all functions from $\mathbb{R}$ to $\mathbb{R}$
+	\item all subsets of $\mathbb{R}$
+\end{itemize}
+
+\section{Example of an Experimental Domain}
+
+Consider the collection of statements, ``this object has a mass strictly between $x$ and $y$ kilograms" where $x$ and $y$ are any positive real numbers with $x<y$.
+Consider the experimental test where we put the object on a balance with an appropriate level of precision for each corresponding statement.
+Then each observation may be equated to an open interval of positive real numbers $(x,y)\subset\mathbb{R}$.
+Conjunction and disjunction of these observations will correspond exactly to intersection and union of these intervals, respectively.
+Further, the collection of intervals of the form $(p,q)$ for $p,q\in\mathbb{Q}$ forms a basis of observations.
+Next, one can show that any sequence of observations satisfying the conditions of exact outcome corresponds precisely to a real number.
+In particular, the set of possibilities $X$ for this experimental domain will be the set of positive real numbers with their usual topology.
+The reader with some experience with elementary real analysis will recognize that this construction is essentially the same as the Cauchy sequence construction of the real numbers, where representative Cauchy sequences come from endpoints of intervals, and the intervals used those corresponding to observations in a sequence of observations yielding an exact outcome.
 
@@
 
+\begin{prop}
+	\label{setfunctions}
+	Let $X$ and $Y$ be two Hausdorff topological spaces, and let $g: \mathsf{T}(Y) \rightarrow \mathsf{T}(X)$ be a mapping such that:
+	\begin{enumerate}
+		\item It is compatible with union and intersection $\forall V_1, V_2 \in Y$ $g(V_1 \cup V_2)=g(V_1)\cup g(V_2)$ and $g(V_1 \cap V_2)=g(V_1)\cap g(V_2)$
+		\item $g(\emptyset) = \emptyset$
+		\item $g(Y) = X$
+	\end{enumerate}
+	Then there exists a unique continuous function $f: X \rightarrow Y$ such that $g = f^{-1} |_{\mathsf{T}(Y)}$.
+\end{prop}
+
+\begin{proof}
+	We claim there exists a unique extension $\bar{g}:\sigma(Y)\to\sigma(X)$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma(X)$ and $\sigma(Y)$, such that $\bar{g}|_{\mathsf{T}(Y)}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
+Let $\bar{g}(V) = g(V)$ for all open sets $V \in \mathsf{T}(Y)$.
+Let $A \in \sigma(Y)$ (not necessarily open) and $A^C$ be its complement.
+We must have $\bar{g}(A^C) = \bar{g}(A)^C = X\setminus \bar{g}(A)$ for $\bar{g}$ to be compatible with complements.
+Recall that all Borel sets in $\sigma(Y)$ and $\sigma(X)$ may be written as some combination of unions, intersections, and complements of open sets.
+Thus, the construction uniquely determine what $\bar{g}$ should output on any Borel set.
+We need only check that the output is still a Borel set.
+But by definition of $\bar{g}$, the outputs will be given as unions, intersections, and complements of outputs of $g$, which are open sets, and so the image of $\bar{g}$ is contained in $\sigma(X)$.
+$\bar{g}$ is well defined.
+The function $\bar{g}$ in a sense represents extracting the maximum amount of information possible out of $g$.
+
+	We claim we can define $\hat{g}:Y\to\sigma(X)$ such that $\hat{g}(y) = \bar{g}(\{y\})$.
+Since $Y$ is Hausdorff, every singleton $\{y\}$ is closed and is therefore a Borel set.
+$\bar{g}(\{y\})$ is well defined and so is $\hat{g}(y)$.
+
+	We claim  $\hat{g}(y_1)\cap\hat{g}(y_2) = \emptyset$ if and only if $y_1\neq y_2$ for all $y_1,y_2\in Y$ such that $\hat{g}(y_i)\neq\emptyset$ for $i=1,2$. If $y_1\neq y_2$ we have
+	$$
+	\hat{g}(y_1)\cap\hat{g}(y_2) = \bar{g}(\{y_1\})\cap\bar{g}(\{y_2\}) = \bar{g}(\{y_1\}\cap\{y_2\}) = \bar{g}(\emptyset) = \emptyset.
+	$$
+	Conversely, if $y_1 = y_2$ we have
+	$$
+	\hat{g}(y_1)\cap\hat{g}(y_2) = 	\hat{g}(y_1)\cap\hat{g}(y_1) = 
+	\hat{g}(y_1) \neq \emptyset.
+	$$
+
+	We claim we can define $f: X\to Y$ such that $f(x) = y$ if and only if $x\in \hat{g}(y)$.
+Since $g(Y)=X$, there exists $y\in Y$ such that $x\in\hat{g}(y)$.
+By the preceding claim, this $y$ is unique.
+$f: X\to Y$ is well defined.
+Note that no arbitrary choice where made so far that lead to the construction of $f$, which is therefore determined uniquely by $g$.
+
+	We claim $g = f^{-1} |_{\mathsf{T}(Y)}$.
+Let $V\in\mathsf{T}(Y)$.
+We want to show $f^{-1}(V) = g(V)$.
+Let $x\in f^{-1}(V)$.
+Then for some $y \in V$ we have $f(x)=y$.
+$x\in \hat{g}(y)$ by construction of $f$.
+$\hat{g}(y) \subset g(V)$ since $\{y\}\subset V$, so $x\in g(V)$.
+$f^{-1}(V) \subseteq g(V)$.
+Conversely, let $x\in g(V)=\bar{g}(V)$.
+Then for some $y\in V$, we have $x\in\bar{g}(\{y\})\subset\bar{g}(V)$.
+But then by definition we have $f(x)=y$, so $x\in f^{-1}(V)$.
+$f^{-1}(V) \supseteq g(V)$.
+$f^{-1}(V) = g(V)$ for all $V\in\mathsf{T}(Y)$ and therefore $f^{-1}|_{\mathsf{T}(Y)}=g$.
+
+	We claim $f$ is continuous. It is so since $g = f^{-1} |_{\mathsf{T}(Y)}$ takes open sets to open sets. 
+\end{proof}
+
+The above work gives us an approach to reconstruct continuous functions given the behavior of the inverse on open sets.
+In the context of collecting information on observable phenomena, the function $g$ represents the total of all information possible to gather on the correlation between two variables.
+The fact that from $g$ we can construct a unique continuous $f$ shows us that in the infinite-resource ideal, we fully obtain the information to give exact relations between variables.
+Further, the result itself tells us that the interesting phenomena in our framework of experimental observations will be continuous, which is often a baseline assumption in physics.
+
+For continuous functions to be physically distinguishable, we need to show they can always be given, as a set, given a topology that is Hausdorff and second-countable.
+
+To that end, we introduce the basis-to-basis topology on the set of continuous functions from two topological spaces.
+This is the topology generated by the sets of functions that map a basis element of one element inside an element of the other space.
+
+\begin{defn} Let $X$ and $Y$ be two topological spaces.
+Let $C(X,Y)$ denote the set of all continuous functions from $X$ to $Y$.
+Let $\mathcal{B}_X$ and $\mathcal{B}_Y$ be two bases for $X$ and $Y$ respectively.
+The basis-to-basis topology $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ on $C(X,Y)$ with respect to the basis $\mathcal{B}_X$ and $\mathcal{B}_Y$ is the topology generated by all sets of the form
+	$$
+	V(U_X, U_Y) = \{f\in C(X,Y) : f(U_X)\subset U_Y\}
+	$$
+where $U_X \in \mathcal{B}_X$ and $U_Y \in \mathcal{B}_Y$.
+\end{defn}
+
+Now we need to show that if the two spaces $X$ and $Y$ are Hausdorff and second-countable, the basis-to-basis topology is Hausdorff and second-countable for a suitable choice of basis.
+
+\begin{prop}
+	Let $X$ and $Y$ be two Hausdorff and second-countable topological spaces.
+Let $C(X,Y)$ denote the set of all continuous functions from $X$ to $Y$.
+Let $\mathcal{B}_X$ and $\mathcal{B}_Y$ be two countable bases for $X$ and $Y$ respectively.
+The basis-to-basis topology $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ on $C(X,Y)$ with respect to the basis $\mathcal{B}_X$ and $\mathcal{B}_Y$ is Hausdorff and second-countable.
+\end{prop}
+\begin{proof}
+	We claim $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is second-countable.
+We first note that the sub-basis $\{V(U_X, U_Y) \, |\,   U_X \in \mathcal{B}_X , U_Y \in \mathcal{B}_Y \}$ is countable since $\mathcal{B}_X$ and $\mathcal{B}_Y$ are countable.
+The collection $\mathcal{B}$ of all finite intersections is still countable, since it is in one-to-one correspondence with the collection of finite subsets of a countable set, which is still countable.
+Therefore $\mathcal{B}$ is a countable basis, which means $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is second-countable.
+
+	We claim $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is Hausdorff.
+Let $f,g:X\to Y$ be distinct continuous functions.
+Then for some $x\in X$, we have $f(x)\neq g(x)$.
+Pick $V_1, V_2$ disjoint open subsets of $Y$ with $f(x)\in V_1$ and $g(x)\in V_2$.
+We may assume (possibly by shrinking $V_1$ or $V_2$) that both are basis elements for the topology of $Y$.
+Let $U=f^{-1}(V_1)\cap g^{-1}(V_2)$.
+Then $U$ is an open neighborhood of $x$.
+We may assume again that $U$ is a basis element for the topology on $X$ by shrinking it if necessary.
+Now, let $T_1$ to be the (sub-)basis element for basis-to-basis topology corresponding to $U$ and $V_1$.
+By construction, $f\in T_1$.
+Similarly, let $T_2$ to be the basis element for the basis-to-basis topology corresponding to $U$ and $V_2$ and containing $g$.
+Since $V_1$ and $V_2$ are disjoint, so are $T_1$ and $T_2$.
+$\mathsf{T}(\mathcal{C})$ is Hausdorff.
+\end{proof}
+
+We will also show that this is not in general equal to the open-open topology.
+
+\begin{prop}
+	In general, the topology $\mathsf{T}(\mathcal{C})$ defined above depends on the bases for $X$ and $Y$. 
+\end{prop}
+\begin{proof}
+	We will give an example of such a case.
+Let $X=Y=\mathbb{R}$ with the usual topology on $\mathbb{R}$.
+Let $\mathcal{B}_1$ be the basis for $\mathbb{R}$ consisting of all open intervals with rational endpoints, and let $\mathcal{B}_2 = \mathcal{B}_1\cup\{(0,\pi)\}$.
+We find a set open in our topology $\mathsf{T}(\mathcal{C})$ using $\mathcal{B}_2$ as our basis which is not open when we use $\mathcal{B}_1$.
+Consider the following set, open in our topology when using $\mathcal{B}_2$:
+	$$
+	V = \{f\in\mathcal{C}| f((0,\pi))\subset(0,1)\}
+	$$
+	This is clearly open because $(0,\pi)$ and $(0,1)$ are basis elements in $\mathcal{B}_2$.
+
+	Next, we show that when using $\mathcal{B}_1$ as our basis for $X=Y=\mathbb{R}$, no finite intersection and/or infinite union of sub-basis elements for $\mathsf{T}(\mathcal{C})$ will equal $V$.
+Henceforth, when we say ``open" subsets, unless otherwise stated, we will refer exclusively to the topology $\mathsf{T}(\mathcal{C})$ when using $\mathcal{B}_1$ for $\mathbb{R}$.
+The smallest sub-basis subsets containing $V$ are those of the form:
+	$$
+	B_r = \{f\in\mathcal{C} | f((0,r))\subset(0,1)\}
+	$$
+	for each $r\in\mathbb{Q}$ with $r<\pi$.
+By ``smallest" we mean any sub-basis elements containing $V$ which are not in the form of $B_r$ will contain some $B_r$.
+Now, any finite intersection of the $B_r$'s will be another $B_r$, but these strictly contain $V$, so we can never obtain $V$ from this sub-basis.
+
+	Similarly, the largest sub-basis sets contained within $V$ are of the form: 
+	$$
+	S_r = \{f\in\mathcal{C} | f((0,r))\subset(0,1)\}
+	$$
+	for each $r\in \mathbb{Q}$ with $r>\pi$.
+Let $S = \cup_{r>\pi}S_r$.
+This is the largest open set in our topology which is contained in $V$.
+Consider the function $f(x) = x/\pi$.
+This is continuous and is an element of $V$.
+But notice $f(\pi)=1$, so $f\notin S_r$ for all $r>\pi$, hence $f\notin S$.
+Thus, $V$ is a set which is open in $\mathsf{T}(\mathcal{C})$ when we use $\mathcal{B}_1$ for $X$ and $Y$, but not open when we use $\mathcal{B}_2$ for $X$ and $Y$.
+\end{proof}
+
+TODO: add corollary environment for this one.
+
+\begin{prop}
+The topology $\mathsf{T}(\mathcal{C})$ on the set of continuous functions $C(\mathbb{R},\mathbb{R})$ from $\mathbb{R}$ to itself is in general not equal to the open-open topology on $C(\mathbb{R},\mathbb{R})$.
+\end{prop}
+\begin{proof}
+In the proof of the previous proposition, note that the basis for the open-open topology is independent of the basis used for the underlying spaces, and in particular is a superset of the basis for $\mathsf{T}(\mathcal{C})$ when using $\mathbb{B}_2$ to construct it.
+Hence the open-open topology is distinct from $\mathsf{T}(\mathcal{C})$ when using the basis $\mathcal{B}_1$ on $\mathbb{R}$.
+\end{proof}
+
+TODO: the above propositions might generalize to any uncountable, second-countable space. This will be considered during the paper-writing process later.
+
+Thus we have shown one can start with two Hausdorff, second-countable spaces, and generate a new Hausdorff second-countable space consisting of all functions between them.
+Topological spaces with these properties make up our class of scientifically interesting spaces, so the space of continuous (distinguishability-preserving) functions between spaces of distinguishable quantities is again a physically distinguishable space.
+Because one can iterate this construction (since the relevant properties are preserved), this is a way to generate arbitrarily many new physically distinguishable spaces encoding important information about the ``lower-order" spaces (i.e.
+they consist of functions which encode a conversion between verifiable sets in different spaces).
+
 \section{Conclusion}
+
+\begin{table}[h]
+	\centering
+\begin{tabular}{p{0.20\textwidth} p{0.7\textwidth}}
+	Math/Topology & Science/Physics \\ 
+	\hline 
+	Hausdorff, second-countable topological space & Experimentally distinguishable space, whose points are the possible values and whose open sets represent the experimentally attainable levels of precision \\
+	Open set & Verifiable set. We can verify experimentally that an element is within the set  \\ 
+	Closed set & Refutable set. We can verify experimentally that an element is not in the set \\ 
+	Basis of a topology & A collection of verifiable sets such that any verifiable set is determined by the basis sets\\
+	Continuous \newline function &  A function between two sets of experimental distinguishable elements that preserves distinguishability \\
+	Homeomorphism &  A perfect equivalence between experimentally distinguishable spaces. \\
+\end{tabular} 
+\caption{Topology to physics dictionary}
+\end{table}
 

### 2017-07-05 Blog/07-WhatIsInformationEntropy.tex
Entropy post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{What is information entropy?}

TL;DR .

Starting from the 1950s, the connection between information theory and thermodynamics has been increasingly made clear. What seems not be as clear is that there is a deep link between information theory and Hamiltonian mechanics as well. I want to explore this link in the next few posts and it is appropriate to first have a clear idea on what information entropy is.

The concept of entropy in physics is unfortunately associated with a certain amount of vagueness. While mathematically is well defined, I have never hear a crisp conceptual understanding (e.g. it measures "disorder", it has to do with how much the system is homogeneous, ...). the conceptual idea conceptually people describe it as disorder, or order, or information, or lack of information or how much the system is homogeneous. Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear physicists defining it as "lack of information" or any other vague nonsense.

The scenario is this: you have a distribution over a set of possibilities. You pick an element according to that distribution. On average, how many bits of information you have to use to tell me which element have you picked? How many one-zeros? If that does make sense yet, it will after we go through a few examples.

\section{Examples}

Suppose you have a bag with 40 balls. 20 are green and 20 are red. Suppose you pick a ball and you want to tell me the color. How many bits of information you need to give me? How many one-zero choices? The answer is one bit: 0 for green and 1 for red. That is: all the information can be encoded into one zero-one choice.

Suppose all 40 balls are green. How many bits of information? The answer is zero bits: I already know you are going to get a green ball.

Suppose there are 10 balls for each of these color: green, red, yellow and blue. How many bits of information? We can set 00 for green, 01 for red, 10 for yellow and 11 for blue. That is two zero-one choices: it's two bits of information.

Shannon's entropy simply generalizes this idea: given a $n$ cases each with a probability $p_i$, the average number of bits required to identify an element is:

\begin{equation}
\label{ShannonEntropy}
H(p_1, ..., p_n) = - \sum_{i=1}^{n} p_i \log p_i
\end{equation}

The base of the logarithm determines the units. If it's $2$, then we are using bits. If it's $e$, the natural log, then we are using nats.

\section{Uniqueness of the expression}

It is important to understand that the above expression for entropy is not arbitrary. Consider the following three requirements.

\begin{enumerate}
	\item $H(p_1, ..., p_n)$ is continuous.
	\item $H(p_1, ..., p_n)$ increases if the number of choice increases. That is, suppose you have two uniform distributions over $n$ and $m$ cases respectively. If $m>n$ then $H(\frac{1}{m}, ..., \frac{1}{m})>H(\frac{1}{n}, ..., \frac{1}{n})$.
	\item It does not change if we group/break down choice. Consider the following trees:
	\begin{figure}[h]
	\includegraphics[scale=0.70]{ShannonTree}
	\centering
	\end{figure}

	We should get the same result whether we have one distribution or we combine the two. That is: $H(\frac{1}{2}, \frac{1}{3}, \frac{1}{6}) = H(\frac{1}{2},\frac{1}{2}) + \frac{1}{2} H(\frac{2}{3}, \frac{1}{3})$. The coefficient $\frac{1}{2}$ appears because the second choice only occurs half the times.
\end{enumerate}

These are basic requirements that an expression that measures information needs to satisfy. It's not just that Shannon's entropy \eqref{ShannonEntropy} satisfies these conditions: it is the only one that does. While we are not going to derive it here, the derivation is not particularly insightful, it is important to understand that it can be derived. Shannon's entropy is not just an expression that we use because it's convenient or because it's known to work: it's the only one that can work.

\section{Relation to physics}

There are a couple of aspect we need to clarify as it allows to understand why such concept is interesting for physics.

First of all, information entropy has nothing to do with knowledge (e.g. what we know about the system or what can be known). This misunderstanding arises because the concept can be applied to probability distribution that represent our knowledge of the system. For example, if there are four boxes and I do not know in which one the ball is hiding, I have a uniform distribution on four cases which has 2 bits of information entropy. Conversely, if I know where it is, the entropy is 0 bits. So, in this and only this case, more entropy correspond to less knowledge.

But not all distributions represent knowledge and/or are statistical. For example, we could have a mass distribution in space. In this case, information entropy does not equate to lack of knowledge. Our knowledge of the mass distribution is perfect: it just happens that the mass is spread over a region. The information entropy of a mass distribution in space is as much a physical property of the system as it is the mass distribution itself and it tells us how many bits are necessary to identify the position of a part of the system. Unlike the previous case, this is objective number. It is this second use that is the most relevant to physics. Nature does not much care about our knowledge: it cares about objective facts.

Another thing to keep in mind is that information entropy is a quantity that relates two levels of description: the distribution and the elements over which the distribution is defined. This is exactly the same for the thermodynamic entropy, which relates the state of the whole system (i.e. the macro-state) with the state of its components (i.e. the micro-states). For Hamiltonian mechanics, we will study the evolution of distribution over position and momentum and see how it relates to the evolution of particles.

\section{Conclusion}

This hopefully helps you understand what information entropy measures: how many bits of informations are required on average to identify an element in a distribution. Not "knowledge" or "lack of information" or "disorder" or any other vague concept. In the end, information entropy is a number: to understand it you need to understand what that number quantifies.

\end{document}

### 2017-07-06 Articles/TopologyAndExperimentalDistinguishability/article.tex
Adding authors
Adding Acknowledgments
Added abstract

@@
 %% Group authors per affiliation:
-\author{Gabriele Carcassi\fnref{myfootnote}}
-\address{Somewhere, Ann Arbor}
-\fntext[myfootnote]{Footnote?}
-
-%% or include affiliations in footnotes:
-\author[mymainaddress,mysecondaryaddress]{University of Michigan}
-\ead[url]{www.umich.edu}
-
-\author[mysecondaryaddress]{Secondary\corref{mycorrespondingauthor}}
-\cortext[mycorrespondingauthor]{Corresponding author}
+\author[umich]{Gabriele Carcassi\corref{cor1}}
 \ead{carcassi@umich.edu}
 
-\address[mymainaddress]{Somewhere else, Ann Arbor}
-\address[mysecondaryaddress]{Somewhere instead, Ann Arbor}
+\author[umich]{Christine A. Aidala}
+\ead{caidala@umich.edu}
+
+\author[umich]{Mark J. Greenfield}
+\ead{markjg@umich.edu}
+
+\address[umich]{University of Michigan, Ann Arbor, MI 48109, USA}
+
+\cortext[cor1]{Corresponding author}
 
 \begin{abstract}
-TODO
+We formalize the relationship between topological spaces and the ability to distinguish objects experimentally, providing understanding and justification as to why topological spaces and continuous functions are pervasive tools in the physical sciences.
+The aim is to use these ideas as a stepping stone to give a more rigorous physical foundation to dynamical systems and, in particular, Hamiltonian dynamics.
+We first define an experimental observation as a statement that can be verified using an experimental procedure.
+We show that observations are not closed under negation and countable conjunction, but are closed under finite conjunction and countable disjunction.
+We then consider observations that identify elements in a set and show how they induce a Hausdorff and second countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
+We then show that only continuous functions preserve experimental distinguishability and that the collection of these functions can be given a Hausdorff and second countable topology.
+This shows that the universe of discourse of experimental distinguishability so defined is closed and consistent.
+
 \end{abstract}
@@
 
+\section*{Acknowledgments}
+Funding for this work was provided in part by the MCubed program of the University of Michigan.
+
 \bibliography{bibliography}

### 2017-07-07 Blog/07-WhatIsInformationEntropy.tex
Update to entropy post

@@
 Starting from the 1950s, the connection between information theory and thermodynamics has been increasingly made clear.
-What seems not be as clear is that there is a deep link between information theory and Hamiltonian mechanics as well.
+What I believe is still largely unnoticed is that there is a deep link between information theory and Hamiltonian mechanics as well.
 I want to explore this link in the next few posts and it is appropriate to first have a clear idea on what information entropy is.
@@
 The concept of entropy in physics is unfortunately associated with a certain amount of vagueness.
-While mathematically is well defined, I have never hear a crisp conceptual understanding (e.g.
+While mathematically is well defined, I have never heard a crisp conceptual understanding (e.g.
 it measures "disorder", it has to do with how much the system is homogeneous, ...).
-the conceptual idea conceptually people describe it as disorder, or order, or information, or lack of information or how much the system is homogeneous.
-Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear physicists defining it as "lack of information" or any other vague nonsense.
+Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear physicists defining it as "lack of information" or any other vague description.
+So, let's see if we do better.
 
@@
 How many one-zeros?
-If that does make sense yet, it will after we go through a few examples.
+How many yes-no answers?
+If that doesn't make sense yet, hopefully it will after we go through a few examples.
 
@@
 20 are green and 20 are red.
-Suppose you pick a ball and you want to tell me the color.
-How many bits of information you need to give me?
-How many one-zero choices?
+Suppose you pick a ball and I want to know the color.
+Suppose I can only ask you yes or no questions.
+How many questions do I have to ask you?
+One suffices: "Is it red?".
+If the answer is yes, I know it's red, if the answer is no, I know it's green.
+We can rephrase by asking: how many one-zero choices?
+That is: how many bits of information you need to give me?
 The answer is one bit: 0 for green and 1 for red.
@@
 
-Suppose all 40 balls are green. How many bits of information? The answer is zero bits: I already know you are going to get a green ball.
+Let's do it again in another case. Suppose all 40 balls are green. How many bits of information? The answer is zero bits: I already know you are going to get a green ball.
 
@@
 
-Shannon's entropy simply generalizes this idea: given a $n$ cases each with a probability $p_i$, the average number of bits required to identify an element is:
+Shannon's entropy simply generalizes this idea: given a $n$ cases each with a probability $p_i$, the average number of bits (i.e. yes/no questions) required to identify an element is:
 
@@
 
-The base of the logarithm determines the units. If it's $2$, then we are using bits. If it's $e$, the natural log, then we are using nats.
+The base of the logarithm determines the units.
+If it's base $2$, then we are using bits.
+But sometimes is useful to use base $e$ (i.e.
+the natural log) since it has nice properties.
+In that case we are using nats, and we have $1$ bit = $\ln 2$ nats.
 
@@
 If $m>n$ then $H(\frac{1}{m}, ..., \frac{1}{m})>H(\frac{1}{n}, ..., \frac{1}{n})$.
-	\item It does not change if we group/break down choice. Consider the following trees:
+	\item $H(p_1, ..., p_n)$ does not change if we group/break down choices. Consider the following trees:
 	\begin{figure}[h]
-	\includegraphics[scale=0.70]{ShannonTree}
+	\includegraphics[scale=0.50]{ShannonTree}
 	\centering
@@
 It's not just that Shannon's entropy \eqref{ShannonEntropy} satisfies these conditions: it is the only one that does.
-While we are not going to derive it here, the derivation is not particularly insightful, it is important to understand that it can be derived.
 Shannon's entropy is not just an expression that we use because it's convenient or because it's known to work: it's the only one that can work.
+
+While we are not going to derive it here, the derivation is not particularly insightful, we can still understand why it is what it is.
+If a particular case is a choice among $n$ equal cases, it's probability is $p_i=\frac{1}{n}$ and the number of yes/no questions we have to make to get to it is $\log_2 n = \log_2 \frac{1}{p_i} = - \log_2 p_i$.
+So, when that particular case is selected, we need $- \log_2 p_i$ bits.
+Now we take the expectation, the average over all cases, and we have $H(p_1, ..., p_n) = - \sum_{i=1}^{n} p_i \log p_i$.
+We now have the average value of yes/no questions (i.e.
+bits) needed to identify an element in the distribution.
 
@@
 
-There are a couple of aspect we need to clarify as it allows to understand why such concept is interesting for physics.
+There are a couple of aspects we need to clarify as it allows to understand why such concept is interesting for physics.
 
@@
 what we know about the system or what can be known).
-This misunderstanding arises because the concept can be applied to probability distribution that represent our knowledge of the system.
+This misunderstanding arises because the concept can be applied to probability distribution that represent our knowledge of the system, and in some cases it may relate to knowledge.
 For example, if there are four boxes and I do not know in which one the ball is hiding, I have a uniform distribution on four cases which has 2 bits of information entropy.
 Conversely, if I know where it is, the entropy is 0 bits.
-So, in this and only this case, more entropy correspond to less knowledge.
+So, in this more entropy correspond to less knowledge.
+But this idea does no generalize.
 
-But not all distributions represent knowledge and/or are statistical.
+Suppose you have equal confidence that you may or may not win a prize.
+That is 50\% chance you get one million dollars and 50\% chance you get nothing.
+That is one bit of information entropy.
+On the other hand, suppose you know you won a prize but you do not know how much.
+50\% chance you won half a million dollars and 50 \% chance you won a million dollars.
+It is still one bit of information entropy.
+Yet, I would argue, you know more in the second case: you know you have half a million dollars you can spend right away.
+
+Moreover, not all distributions represent knowledge and/or are statistical.
 For example, we could have a mass distribution in space.
 In this case, information entropy does not equate to lack of knowledge.
-Our knowledge of the mass distribution is perfect: it just happens that the mass is spread over a region.
+Our knowledge of the mass distribution is perfect: we know exactly how it is spread over a region.
 The information entropy of a mass distribution in space is as much a physical property of the system as it is the mass distribution itself and it tells us how many bits are necessary to identify the position of a part of the system.
-Unlike the previous case, this is objective number.
+Unlike the previous case, this is an objective number.
 It is this second use that is the most relevant to physics.
@@
 Another thing to keep in mind is that information entropy is a quantity that relates two levels of description: the distribution and the elements over which the distribution is defined.
+It is this property that makes it important in physics.
 This is exactly the same for the thermodynamic entropy, which relates the state of the whole system (i.e.
@@
 the micro-states).
-For Hamiltonian mechanics, we will study the evolution of distribution over position and momentum and see how it relates to the evolution of particles.
+For Hamiltonian mechanics, we will study the evolution of distributions over position and momentum and see how it relates to the evolution of particles.
+Note that other types of entropy (e.g.
+topological entropy, algebraic entropy) do not have this property and still do not quite understand what they actually quantify.
 

### 2017-07-10 Blog/07-WhatIsInformationEntropy.tex
Started another post on entropy

@@
 
-TL;DR .
+TL;DR Information entropy tells you the average number of bits (i.e. answers to yes/no questions) you need to identify an element from a distribution.
 
@@
 What I believe is still largely unnoticed is that there is a deep link between information theory and Hamiltonian mechanics as well.
-I want to explore this link in the next few posts and it is appropriate to first have a clear idea on what information entropy is.
+I want to explore this link in the next few posts so I'd like to first give a clear idea on what information entropy actually is.
 
@@
 Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear physicists defining it as "lack of information" or any other vague description.
-So, let's see if we do better.
+So, let's see if we can do better.
 
@@
 How many questions do I have to ask you?
-One suffices: "Is it red?".
+One suffices: "is it red?".
 If the answer is yes, I know it's red, if the answer is no, I know it's green.

### 2017-07-10 Blog/08-HamiltonianMechanicsIsConservationOfInformationEntropy.tex
Started another post on entropy

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Hamiltonian mechanics is conservation of information entropy}

TL;DR Hamiltonian systems are those that conserve information entropy during time evolution.

This is one of those result that I figured out a few years ago but I still don't know if and where to publish it. Fortunately for me, I care a lot more about figuring things out than publishing papers.

The idea is the following: suppose you have a distribution over position and momentum $\rho(x, p)$. Suppose you evolve it in time and get a final distribution $\hat{\rho}(x, p)$ using Hamiltonian evolution. That is: you take each little volume and move it according to the trajectories given by Hamiltonian mechanics. How does the information entropy change? Answer: it doesn't. Suppose you look for all possible time evolutions that conserve information entropy, what do you get? You get Hamiltonian mechanics.

That is: Hamiltonian systems are the ones and only the ones that preserve information entropy of a distribution $\rho(x, p)$ over states. That is: identifying an element in the initial distribution requires exactly the same information as an element in the final distribution. That is: knowing the initial element you know the final element and vice-versa. That is: the evolution is deterministic and reversible.

Let's see how the math works.

\section{Hamiltonian mechanics conserves information entropy}

Suppose we have a distribution over position and momentum $\rho(x, p)$. How does it change during Hamiltonian evolution? Suppose we start with a little region $dx \,dp$. The fraction contained in that region will be $\rho(x, p) dx \,dp$. After an infinitesimal time the new coordinates, according to Hamiltonian evolution, will be:
\begin{equation}
\begin{aligned}
\hat{x} = x + \frac{dx}{dt} dt = x + \frac{\partial H}{\partial p} dt  \\
\hat{p} = p + \frac{dp}{dt} dt = p - \frac{\partial H}{\partial x} dt  \\
\end{aligned}
\label{newCoordinates}
\end{equation}

In general, the density and the volume would change according to the Jacobian $|J|$ of the transformation:
\begin{equation}
\begin{aligned}
|J| &= \frac{\partial \hat{x}}{\partial x} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial x} \frac{\partial \hat{x}}{\partial p}\\
d\hat{x}\,d\hat{p} &= |J| dx \,dp  \\
\hat{\rho}(\hat{x}, \hat{p}) &= \frac{\rho(x, p)}{|J|}  \\
\end{aligned}
\label{newDistribution}
\end{equation}
If you are familiar with the Poisson bracket, you can see that the Jacobian is simply the Poisson bracket of the new position and momentum.

Because the evolution is Hamiltonian, the Liouville's theorem applies so the Jacobian is unitary: the volume $d\hat{x}\,d\hat{p} = dx \,dp$ remains unchanged and the final density at the final state is equal to the initial density at the initial state: $\rho(\hat{x}, \hat{p}) = \rho(x, p)$.

How does the information entropy change? We calculate the entropy at the final state and then change variable using \eqref{newDistribution}:
\begin{equation}
\begin{aligned}
-\int \hat{\rho} \log \hat{\rho} \; d\hat{x}\,d\hat{p}&= -\int \frac{\rho}{|J|} \log \frac{\rho}{|J|} \; |J| dx \,dp = -\int \frac{\rho}{|J|} \log \frac{\rho}{|J|} \; |J| dx \,dp \\
&= -\int \rho \log \frac{\rho}{|J|} \; dx \,dp \\
&= -\int \rho \log \rho \; dx \,dp + \int \rho \log |J| \; dx \,dp \\
\end{aligned}
\label{newEntropy}
\end{equation}
That is: the information entropy of the final distribution is equal to the one of the initial distribution plus the expectation of the Jacobian.

But since the evolution is Hamiltonian, the Jacobian is unitary. The information entropy of the final distribution is that of the initial distribution. The information entropy is conserved.

\section{Conservation of information entropy gives Hamiltonian mechanics}

\end{document}

### 2017-07-11 Articles/TopologyAndExperimentalDistinguishability/article.tex
Adding introduction, conclusion and introductions for sections

@@
 
+The successful use of mathematical ideas in experimental sciences is something long established and celebrated.\footnote{Unreasonable effectiveness and other} In particular topology is arguably the most widespread, as it provides the foundation for other mathematical tools such as differential geometry and ???.
+Why is it so successful?
+What physical property is captured by topological spaces that is so fundamental for scientific investigation?
+
+We believe we have identified the answer: topologies capture the way objects are experimentally distinguishable.
+It is open sets that are associated with experimental outcomes, not the values themselves.
+For example, as real values are only measurable with finite precisions, the standard topology on euclidean spaces is the one that is physically meaningful.
+Along the same line, experimental relationships will need to preserve what can be experimentally distinguished and they will therefore be continuous functions.
+
+In this sense, experimentally distinguishability is \emph{the} application of topology in experimental science.
+It is fitting, then, that topology is so wide-spread as defining what can be experimentally identified is truly a fundamental aspect in science.
+The aim of this paper is to lay down a formal framework that formalizes this insight.
+
+The work is organized as follows.
+We will first define experimental observations as statements paired with an experimental test that is able to verify them.
+We will study their properties under logical operations and show they are only closed under finite conjunction and countable disjunction.
+
+We will then show that any set of objects that are experimentally distinguishable is a Hausdorff second-countable topological space.
+Both the points and the topology of the space will be formally defined in terms of the observations themselves.
+
+Finally, we will study relationships between experimentally distinguishable objects. This will be represented by continuous functions and can be shown to be themselves physically distinguishable
+
+We believe this work not only provides interest insights, but a firm justification of the use of topological ideas in the physical science.
+
 \section{Experimental observations}
+
+In science, something is accepted as true only if one can provide a way to independently test the assertion. We introduce the following definitions to capture this notion.
 
@@
 Therefore it is improper to say that one observation verifies another: only experimental tests verify observations.
+
+\section{Algebra of experimental observations}
+
+We now want to understand how experimental observations behave under logical operations. The resulting algebra will have some key differences from the standard Boolean algebra.
 
@@
 
+Being able to test experimental observations one at a time is not enough.
+Given a set of observations we must be able to test them all to find those that are verified.
+With this in mind, we introduce the following definitions.
+
 \begin{defn}
@@
 \section{Experimental distinguishability}
+
+Given an experimental domain, we want to study the maximal sets of consistent observations that can be verified at the same time.
+These will define the possibilities of our domain: the possible cases that can be experimentally distinguished through the observations.
 
@@
 
+Now that we have defined the possibilities of our domains and the topology associated with them, we want to establish relationships between them.
+We want to do this because the aim of scientific investigation is studying and characterizing these relationships
+
 \begin{prop}
@@
 
+We have shown that it is possible to formally capture the requirements of experimental sciences using experimental observations and their properties.
+These definitions lead naturally to topological spaces giving the result that any set of experimentally distinguishable possibilities is a Hausdorf and second-countable topological space.
+In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects as they too can be given a Hausdorf and secound-countable topology.
+The ability to create functions of any order confirms that the universe of discourse provided by these definition is closed and self-consistent.
+
+We hope that this work, because of both its method and result, can provide a more solid foundation to formalize experimental sciences.
+
 \begin{table}[h]

### 2017-07-11 Blog/07-WhatIsInformationEntropy.tex
Final pass at first post on entropy

@@
 it measures "disorder", it has to do with how much the system is homogeneous, ...).
-Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear physicists defining it as "lack of information" or any other vague description.
+Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear some defining it as "lack of information" or any other vague description.
 So, let's see if we can do better.
@@
 
-Shannon's entropy simply generalizes this idea: given a $n$ cases each with a probability $p_i$, the average number of bits (i.e. yes/no questions) required to identify an element is:
+Shannon's entropy simply generalizes this idea: given a $n$ cases each with a probability $p_i$, the average number of bits (i.e. yes/no answers) required to identify an element is:
 
@@
 
-There are a couple of aspects we need to clarify as it allows to understand why such concept is interesting for physics.
+There are a couple of aspects we need to clarify as it allows to understand why such concept is useful in physics.
 
@@
 Conversely, if I know where it is, the entropy is 0 bits.
-So, in this more entropy correspond to less knowledge.
+So, in this case more entropy correspond to less knowledge.
 But this idea does no generalize.
@@
 Suppose you have equal confidence that you may or may not win a prize.
-That is 50\% chance you get one million dollars and 50\% chance you get nothing.
-That is one bit of information entropy.
+That is: 50\% chance you get one million dollars and 50\% chance you get nothing.
+That corresponds to one bit of information entropy.
 On the other hand, suppose you know you won a prize but you do not know how much.
-50\% chance you won half a million dollars and 50 \% chance you won a million dollars.
+50\% chance you won half a million dollars and 50\% chance you won a million dollars.
 It is still one bit of information entropy.
@@
 Note that other types of entropy (e.g.
-topological entropy, algebraic entropy) do not have this property and still do not quite understand what they actually quantify.
+topological entropy, algebraic entropy) do not have this property and I still do not quite understand what they actually quantify.
 
@@
 Not "knowledge" or "lack of information" or "disorder" or any other vague concept.
-In the end, information entropy is a number: to understand it you need to understand what that number quantifies.
+In the end, information entropy is a number: to understand it you need to understand what that number quantifies and in what units.
 

### 2017-07-11 Blog/08-HamiltonianMechanicsIsConservationOfInformationEntropy.tex
Finished second paper on entropy

@@
 How does it change during Hamiltonian evolution?
-Suppose we start with a little region $dx \,dp$.
-The fraction contained in that region will be $\rho(x, p) dx \,dp$.
+Start with a little region $dx \,dp$.
+The fraction of the distribution contained in that region will be $\rho(x, p) dx \,dp$.
 After an infinitesimal time the new coordinates, according to Hamiltonian evolution, will be:
@@
 \begin{aligned}
-|J| &= \frac{\partial \hat{x}}{\partial x} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial x} \frac{\partial \hat{x}}{\partial p}\\
+|J| &= \left| \begin{matrix}
+\dfrac{\partial \hat{x}}{\partial x} & \dfrac{\partial \hat{x}}{\partial p} \\[2.2ex]
+\dfrac{\partial \hat{p}}{\partial x} & \dfrac{\partial \hat{p}}{\partial p} \end{matrix} \right| = \frac{\partial \hat{x}}{\partial x} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial x} \frac{\partial \hat{x}}{\partial p}\\
 d\hat{x}\,d\hat{p} &= |J| dx \,dp  \\
@@
 \end{equation}
-If you are familiar with the Poisson bracket, you can see that the Jacobian is simply the Poisson bracket of the new position and momentum.
+BTW: if you are familiar with the Poisson bracket, you can see that the Jacobian is simply the Poisson bracket of the new position and momentum. But that's for another time.
 
-Because the evolution is Hamiltonian, the Liouville's theorem applies so the Jacobian is unitary: the volume $d\hat{x}\,d\hat{p} = dx \,dp$ remains unchanged and the final density at the final state is equal to the initial density at the initial state: $\rho(\hat{x}, \hat{p}) = \rho(x, p)$.
+Because the evolution is Hamiltonian, the Liouville's theorem applies so the Jacobian is unitary. Let's verify that:
+\begin{equation}
+\begin{aligned}
+|J| &= \left| \begin{matrix}
+1 + \dfrac{\partial}{\partial x} \left( \dfrac{dx}{dt} \right) dt & \dfrac{\partial}{\partial p} \left( \dfrac{dx}{dt} \right) dt \\[2.2ex]
+\dfrac{\partial}{\partial x} \left( \dfrac{dp}{dt} \right) dt & 1 + \dfrac{\partial}{\partial p} \left( \dfrac{dp}{dt} \right) dt \end{matrix} \right| \\
+&= 1 + \left[ \dfrac{\partial}{\partial x} \left( \dfrac{dx}{dt} \right) + \dfrac{\partial}{\partial p} \left( \dfrac{dp}{dt} \right) \right] dt + O(dt^2)\\
+&= 1 + \left[ \dfrac{\partial}{\partial x} \frac{\partial H}{\partial p} - \dfrac{\partial}{\partial p} \frac{\partial H}{\partial x} \right] dt + O(dt^2)\\
+&= 1 + O(dt^2)\\
+\end{aligned}
+\label{Jacobian}
+\end{equation}
+
+This means the volume $d\hat{x}\,d\hat{p} = dx \,dp$ remains unchanged and the final density at the final state is equal to the initial density at the initial state: $\rho(\hat{x}, \hat{p}) = \rho(x, p)$.
 
@@
 
+We can run the argument backwards.
+If information entropy is conserved, the logarithm of the Jacobian has to be zero everywhere, which can only happen if the Jacobian is equal to one everywhere.
+Looking at \eqref{Jacobian}, this happens only if:
+\begin{equation}
+\dfrac{\partial}{\partial x} \left( \dfrac{dx}{dt} \right) + \dfrac{\partial}{\partial p} \left( \dfrac{dp}{dt} \right) = 0
+\label{Jacobian}
+\end{equation}
+
+This means that the vector field formed by the displacement vector $\left[ \dfrac{dx}{dt}, \dfrac{dp}{dt}\right]$ is divergence free. Which means it admits a potential $H$ such that:
+\begin{equation}
+\begin{aligned}
+\frac{dx}{dt} &= \frac{\partial H}{\partial p}  \\
+\frac{dp}{dt} &= - \frac{\partial H}{\partial x}
+\end{aligned}
+\label{Hamilton}
+\end{equation}
+So we have recovered Hamilton's equations.
+
+\section{Conclusion}
+As we have saw, Hamiltonian mechanics is equivalent to conservation of information entropy.
+This idea can be generalized to multiple independent degrees of freedoms.
+The math is a bit more involved but does not add to the general idea.
+
+This is why isolated systems conserve energy: if they are isolated it means that their motion only depends on the information that is already contained in the system.
+Which means it cannot change.
+Which gives us Hamiltonian motion.
+Which means: energy is conserved.
+
 \end{document}

### 2017-07-12 Articles/TopologyAndExperimentalDistinguishability/article.tex
Fixed author formatting
Shortened abstract

@@
 %% Group authors per affiliation:
-\author[umich]{Gabriele Carcassi\corref{cor1}}
+\author[]{Gabriele Carcassi\corref{cor1}}
 \ead{carcassi@umich.edu}
 
-\author[umich]{Christine A. Aidala}
+\author[]{Christine A. Aidala}
 \ead{caidala@umich.edu}
 
-\author[umich]{Mark J. Greenfield}
+\author[]{Mark J. Greenfield}
 \ead{markjg@umich.edu}
 
-\address[umich]{University of Michigan, Ann Arbor, MI 48109, USA}
+\address{University of Michigan, Ann Arbor, MI 48109, USA}
 
@@
 The aim is to use these ideas as a stepping stone to give a more rigorous physical foundation to dynamical systems and, in particular, Hamiltonian dynamics.
-We first define an experimental observation as a statement that can be verified using an experimental procedure.
-We show that observations are not closed under negation and countable conjunction, but are closed under finite conjunction and countable disjunction.
-We then consider observations that identify elements in a set and show how they induce a Hausdorff and second countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
-We then show that only continuous functions preserve experimental distinguishability and that the collection of these functions can be given a Hausdorff and second countable topology.
-This shows that the universe of discourse of experimental distinguishability so defined is closed and consistent.
+We first define an experimental observation as a statement that can be verified using an experimental procedure and show that observations are closed under finite conjunction and countable disjunction.
+We then consider observations that identify elements in a set and show how they induce a Hausdorff and second-countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
+We then show that only continuous functions preserve experimental distinguishability and that the collection of these functions can be given a Hausdorff and second-countable topology.
 

### 2017-07-12 Articles/TopologyAndExperimentalDistinguishability/article.tex
Minor edits of Introduction

@@
 It is open sets that are associated with experimental outcomes, not the values themselves.
-For example, as real values are only measurable with finite precisions, the standard topology on euclidean spaces is the one that is physically meaningful.
-Along the same line, experimental relationships will need to preserve what can be experimentally distinguished and they will therefore be continuous functions.
+For example, as real values are only measurable with finite precision, the standard topology on Euclidean spaces is the one that is physically meaningful.
+Along the same line, experimental relationships need to preserve what can be experimentally distinguished, and they are therefore represented as continuous functions.
 
-In this sense, experimentally distinguishability is \emph{the} application of topology in experimental science.
-It is fitting, then, that topology is so wide-spread as defining what can be experimentally identified is truly a fundamental aspect in science.
-The aim of this paper is to lay down a formal framework that formalizes this insight.
+In this sense, experimental distinguishability is \emph{the} application of topology in experimental science.
+It is fitting, then, that topology is so widespread given that defining what can be experimentally identified is truly a fundamental aspect of science.
+The aim of this paper is to lay down a framework that formalizes this insight.
 
 The work is organized as follows.
-We will first define experimental observations as statements paired with an experimental test that is able to verify them.
-We will study their properties under logical operations and show they are only closed under finite conjunction and countable disjunction.
+We first define experimental observations as statements paired with an experimental test that is able to verify them.
+We study their properties under logical operations and show that they are only closed under finite conjunction and countable disjunction.
+We then show that any set of objects that are experimentally distinguishable is a Hausdorff second-countable topological space.
+Both the points and the topology of the space are formally defined in terms of the observations themselves.
+Finally, we study relationships between experimentally distinguishable objects.
+These are represented by continuous functions, and they can be shown to be themselves physically distinguishable.
 
-We will then show that any set of objects that are experimentally distinguishable is a Hausdorff second-countable topological space.
-Both the points and the topology of the space will be formally defined in terms of the observations themselves.
-
-Finally, we will study relationships between experimentally distinguishable objects. This will be represented by continuous functions and can be shown to be themselves physically distinguishable
-
-We believe this work not only provides interest insights, but a firm justification of the use of topological ideas in the physical science.
+We believe this work not only provides interesting insights, but a firm justification of the use of topological ideas in the physical sciences.
 

### 2017-07-12 Articles/TopologyAndExperimentalDistinguishability/article.tex
Fix typos in Conclusion

@@
 We have shown that it is possible to formally capture the requirements of experimental sciences using experimental observations and their properties.
-These definitions lead naturally to topological spaces giving the result that any set of experimentally distinguishable possibilities is a Hausdorf and second-countable topological space.
-In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects as they too can be given a Hausdorf and secound-countable topology.
-The ability to create functions of any order confirms that the universe of discourse provided by these definition is closed and self-consistent.
+These definitions lead naturally to topological spaces, giving the result that any set of experimentally distinguishable possibilities is a Hausdorff and second-countable topological space.
+In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects, as they too can be given a Hausdorff and secound-countable topology.
+The ability to create functions of any order confirms that the universe of discourse provided by these definitions is closed and self-consistent.
 

### 2017-07-13 Articles/TopologyAndExperimentalDistinguishability/article.tex
started with new definition of possibility, ran into problems

--- 2017-07-12 Articles/TopologyAndExperimentalDistinguishability/article.tex
@@
 \usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd} 
+\usepackage{MnSymbol,centernot}
+
+% Symbol for "implies" and "not implies"
+\def\imp{\Rightarrow}
+\def\nimp{\nRightarrow}
+
+% Symbol for "compatibility" and "incompatibility"
+\def\comp{\doublefrown}
+\def\ncomp{\ndoublefrown}
 
@@
 %% Group authors per affiliation:
-\author[]{Gabriele Carcassi\corref{cor1}}
+
+\author[umich]{Christine A. Aidala}
+\ead{caidala@umich.edu}
+
+\author[umich]{Gabriele Carcassi\corref{cor1}}
 \ead{carcassi@umich.edu}
 
-\author[]{Christine A. Aidala}
-\ead{caidala@umich.edu}
-
-\author[]{Mark J. Greenfield}
+\author[umich]{Mark J. Greenfield}
 \ead{markjg@umich.edu}
 
-\address{University of Michigan, Ann Arbor, MI 48109, USA}
+\address[umich]{University of Michigan, Ann Arbor, MI 48109, USA}
 
@@
 The aim is to use these ideas as a stepping stone to give a more rigorous physical foundation to dynamical systems and, in particular, Hamiltonian dynamics.
-We first define an experimental observation as a statement that can be verified using an experimental procedure and show that observations are closed under finite conjunction and countable disjunction.
-We then consider observations that identify elements in a set and show how they induce a Hausdorff and second-countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
-We then show that only continuous functions preserve experimental distinguishability and that the collection of these functions can be given a Hausdorff and second-countable topology.
+We first define an experimental observation as a statement that can be verified using an experimental procedure.
+We show that observations are not closed under negation and countable conjunction, but are closed under finite conjunction and countable disjunction.
+We then consider observations that identify elements in a set and show how they induce a Hausdorff and second countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
+We then show that only continuous functions preserve experimental distinguishability and that the collection of these functions can be given a Hausdorff and second countable topology.
+This shows that the universe of discourse of experimental distinguishability so defined is closed and consistent.
 
@@
 \begin{defn}
-	A \textbf{statement} $\mathsf{s}$ is a declarative sentence that is either true or false, as in classical logic.
+	A \emph{statement} $\mathsf{s}$ is a declarative sentence that is either true or false, as in classical logic. Here, we allow for infinite statements. 
+\end{defn}
+
+In addition to the usual logical operations on statements, we introduce the following notation:
+
+\begin{defn}
+We say two statements $s$ and $t$ are \emph{compatible}, denoted $s\comp t$, if the conjunction $s\wedge t$ is not a contradiction; otherwise they are \emph{incompatible}, denoted $s\ncomp t$. 
+\end{defn}
+
+Given a statement, we wish to consider when it makes scentific sense to test. The following construction will define the building blocks of this theory.
+
+\begin{defn}
+	An \emph{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e. it can be restarted and stopped an arbitrary number of times) that, if successful, terminates in a finite amount of time.
 \end{defn}
@@
 \begin{defn}
-	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure that can terminate successfully in a finite amount of time and that can be stopped at any time.
-\end{defn}
-
-TODO: in principle, time can be defined as a monoid. This would prevent circular arguments to some extent.
-
-\begin{defn}
-	An \textbf{experimental observation} is a tuple $\mathsf{o} = \llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is always successful.
-The experimental observation is \textbf{verified} if the statement is true.
+	An \emph{experimental observation} is a tuple $\mathsf{o} = \llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is always successful.
+The experimental observation is \emph{verified} if the statement is true.
 \end{defn}
@@
 \begin{defn}
-	An \textbf{experimental counter-observation} is a tuple $\mathsf{o}^c= \llbracket \mathsf{s}, \mathsf{e} \rrbracket$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is false if and only if the experimental test is always successful.
-The experimental counter-observation is \textbf{refuted} if the statement is false.
+	An \emph{experimental counter-observation} is a tuple $\mathsf{o}^c= \llbracket \mathsf{s}, \mathsf{e} \rrbracket$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is false if and only if the experimental test is always successful.
+The experimental counter-observation is \emph{refuted} if the statement is false.
 \end{defn}
 
-We also define some basic relationships between experimental observations.
+Experimental observations are the core object of the scientific discourse. We now define some basic relationships between experimental observations.
 
 \begin{defn}
-	Given two experimental observations $\mathsf{o}_1$ and $\mathsf{o}_2$:
+	Given two experimental observations $\mathsf{o}_1$ and $\mathsf{o}_2$ we define:
 	\begin{itemize}
-	\item \textbf{implication}: $\mathsf{o}_1 \implies \mathsf{o}_2$ if $\mathsf{o}_2$ is verified whenever $\mathsf{o}_1$ is verified
-	\item \textbf{equivalence}: $\mathsf{o}_1 = \mathsf{o}_2$ if $\mathsf{o}_1 \implies \mathsf{o}_2$ and $\mathsf{o}_2 \implies \mathsf{o}_1$
-	\item \textbf{mutual exclusion}: $\mathsf{o}_1 \asymp \mathsf{o}_2$ ($\mathsf{o}_1$ excludes $\mathsf{o}_2$) if one is not verified whenever the other one is
+	\item \emph{implication}: $\mathsf{o}_1 \imp \mathsf{o}_2$ ($\mathsf{o}_1$ implies $\mathsf{o}_2$) if $\mathsf{o}_2$ is always verified whenever $\mathsf{o}_1$ is verified
+	\item \emph{equivalence}: $\mathsf{o}_1 = \mathsf{o}_2$ ($\mathsf{o}_1$ is equal to $\mathsf{o}_2$) if $\mathsf{o}_1 \imp \mathsf{o}_2$ and $\mathsf{o}_2 \imp \mathsf{o}_1$
+	\item \emph{compatibility}: $\mathsf{o}_1 \comp \mathsf{o}_2$ ($\mathsf{o}_1$ is compatible with $\mathsf{o}_2$) if there exists a case in which $\mathsf{o}_1$ and $\mathsf{o}_2$ are jointly verified
 	\end{itemize}
+    The negation of these properties will be noted by $\nimp$, $\neq$ , $\ncomp$ respectively.
 \end{defn}
-TODO: add a symbol for ``not mutually exclusive". I couldn't find the $\asymp$ with a slash through it.
 
@@
 Therefore it is improper to say that one observation verifies another: only experimental tests verify observations.
+
+Lastly, we define a special type of observation.
+
+\begin{defn}
+	The \emph{null observation} is any experimental observation that is never verified (i.e. the experimental test will never terminate successfully). 
+\end{defn}
+
+The null observation is the degenerate case and has the following, maybe counterintuitive, properties.
+Since it is never verified, it can never rule out any other observations: the null observation implies all other observations.
+Since it is never verified, it can never be jointly verified with another observation: the null observation is incompatible with every experimental observation, including itself.
 
@@
 We define the following logical operations on experimental observations.
-Taken as whole, they define the \textbf{algebra of experimental observations}.
+Taken as whole, they define the \emph{algebra of experimental observations}.
 With each definition, we will prove that the resulting object is an experimental (counter-)observation.
@@
 \begin{defn}
-	The \textbf{negation or logical NOT} of an experimental observation $\mathsf{o}=\llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ is the experimental counter-observation $\neg \mathsf{o}=\llbracket\neg \mathsf{s}, \mathsf{e}\rrbracket$ where the negation of the statement $\mathsf{s}$ can be refuted by the experimental test $\mathsf{e}$.
+	The \emph{negation} or \emph{logical NOT} of an experimental observation $\mathsf{o}=\llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ is the experimental counter-observation $\neg \mathsf{o}=\llbracket\neg \mathsf{s}, \mathsf{e}\rrbracket$ where the negation of the statement $\mathsf{s}$ can be refuted by the experimental test $\mathsf{e}$.
 \end{defn}
@@
 \begin{defn}
-	The \textbf{conjunction or logical AND} of a finite collection  of experimental observations $\{\mathsf{o}_i\}_{i=1}^{n}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{n}$ is the experimental observation $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigwedge\limits_{i=1}^{n} \mathsf{s}_i$ is the conjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ is the experimental test that successfully terminates if and only if all $\{\mathsf{e}_i\}_{i=1}^{n}$ successfully terminate.
+	The \emph{conjunction} or \emph{logical AND} of a finite collection  of experimental observations $\{\mathsf{o}_i\}_{i=1}^{n}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{n}$ is the experimental observation $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigwedge\limits_{i=1}^{n} \mathsf{s}_i$ is the conjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ is the experimental test that successfully terminates if and only if all $\{\mathsf{e}_i\}_{i=1}^{n}$ successfully terminate.
 \end{defn}
@@
 \begin{defn}
-	The \textbf{disjunction or logical OR} of a countable (finite or infinite) collection of experimental observations $\{\mathsf{o}_i\}_{i=1}^{\infty}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{\infty}$ is the experimental observation $\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigvee\limits_{i=1}^{\infty} \mathsf{s}_i$ is the disjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ is the experimental test that successfully terminates if and only if at least one experimental test in $\{\mathsf{e}_i\}_{i=1}^{\infty}$ successfully terminates.
+	The \emph{disjunction} or \emph{logical OR} of a countable (finite or infinite) collection of experimental observations $\{\mathsf{o}_i\}_{i=1}^{\infty}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{\infty}$ is the experimental observation $\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigvee\limits_{i=1}^{\infty} \mathsf{s}_i$ is the disjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ is the experimental test that successfully terminates if and only if at least one experimental test in $\{\mathsf{e}_i\}_{i=1}^{\infty}$ successfully terminates.
 \end{defn}
@@
 
-For technical reasons, we will need the following:
-
-\begin{defn}
-The \textbf{empty observation} is any experimental observation of the form $\llparenthesis \mathsf{s}\wedge \neg\mathsf{s},\mathsf{e}\rrparenthesis$. Note that this will always fail. 
-\end{defn}
-
 \section{Experimental domain}
@@
 \begin{defn}
-	An \textbf{generalized experimental domain} is a set of observations closed under finite conjunction and countable disjunction.
-Any subset that can generate all others via finite conjunction and countable disjunction is called a \textbf{sub-basis} of the domain.
-A \textbf{basis} is a subset of the observations which can generate all others by countable disjunctions.
+	An \emph{experimental domain} is a set of observations closed under finite conjunction and countable disjunction such that all observations can be tested in infinite time. 
+\end{defn}
+
+\begin{defn}
+	A \emph{sub-basis} of an experimental domain is any subset that can generate all others via finite conjunction and countable disjunction.
+A \emph{basis} of an experimental domain is any subset that can generate all others by countable disjunctions.
 \end{defn}
@@
 \begin{prop}
-Let $\mathcal{S}$ be an experimental domain.
-Then it is possible to verify all experimental observations $\mathsf{o}\in S$ in infinite time if and only if there exists a countable basis (equivalently, sub-basis) $\mathcal{B}$ of $\mathcal{S}$.
+Let $\mathcal{S}$ be an experimental domain. Then there exists a countable basis (equivalently, sub-basis) $\mathcal{B}$ of $\mathcal{S}$.
 \end{prop}
+
 \begin{proof}
@@
 If there does not exist a countable basis, then by definition there does not exist a sequence of experimental observations in $\mathcal{S}$ from which one can deduce all other observations in $\mathcal{S}$.
-Hence it is impossible to verify all members of $\mathcal{S}$.
+Hence it is impossible to test all members of $\mathcal{S}$.
 \end{proof}
 
-In light of the above result, we have the following definition, on which the remainder of this chapter will focus.
-
-\begin{defn}
-An \textbf{experimental domain} is a set of observations closed under finite conjunction and countable disjunction which has a countable basis and contains the empty observation. 
-\end{defn}
+\begin{rem} 
+It is not hard to see that the (countable) disjunction of an entire basis produces an experimental observation which is always verified. Henceforth, we call this the \emph{trivial observation}. 
+\end{rem}
 
@@
 We will begin with a given experimental domain $\mathcal{S}$.
+First, we will construct an intermediate space which will mimic the Borel algebra of a topological space:
 
 \begin{defn}
-A \textbf{generalized exact outcome} for $\mathcal{S}$ is a sequence of experimental observations $\mathcal{P} = \{\mathsf{o}_i\}_{i=1}^{\infty}\subset\mathcal{S}$ with the following property.
-For all experimental observations $\mathsf{o}\in\mathcal{S}$, there exists some positive integer $N$ such that for all integers $n>N$, one of the following must happen:
-\begin{enumerate}
-\item $\mathsf{o}_n \asymp \mathsf{o}$ 
-\item $\mathsf{o}_n \implies \mathsf{o}$
-\end{enumerate}
+Given an experimental domain $\mathcal{S}$, define the associated \emph{theoretical domain} $\mathcal{S}^*$ to be the closure under countable conjunction, disjunction, and negation of the set of statements in all experimental observations of $\mathcal{S}$.
 \end{defn}
 
-\begin{defn}
-An \textbf{exact outcome} for $\mathcal{S}$ is a generalized exact outcome $\mathcal{P} = \{\mathsf{o}_i\}_{i=1}^{\infty}$ such that for all $i\geq1$, we have $\mathsf{o}_{i+1}\implies\mathsf{o}_i$.
-\end{defn}
-Note that given a generalized exact outcome, one can build an exact outcome by simply replacing each $\mathsf{o}_n$ with the (finite) conjunction $\bigvee_{i=1}^n\mathsf{o}_i$.
-We will often denote the collection of (generalized) exact outcomes by $\mathcal{C}$.
-Next, we note that different sequences can result in ``equivalent" outcomes.
+Next, we will extract the appropriate subcollection of the theoretical domain in order to obtain what will serve as ``points" in our space:
 
 \begin{defn}
-We say two exact outcomes $\mathcal{P} = \{\mathsf{o}_i\}_{i=1}^{\infty}$ and $\mathcal{P}' = \{\mathsf{o}'_i\}_{j=1}^{\infty}$ are \textbf{equivalent}, denoted $\mathcal{P}\sim\mathcal{P}'$, if for all $i,j$, we have that $\mathsf{o}_i$ and $\mathsf{o}'_j$ are not mutually exclusive.
+\label{possibilitiesdef}
+Given a theoretical domain $\mathcal{S}^*$, define the set $X$ of \emph{possibilities} to be the following subset:
+$$
+X = \{s\in\mathcal{S}^*\ |\ \forall t\in\mathcal{S}^*, \text{ either } s\imp t \text{ or } s\ncomp t\}
+$$
 \end{defn}
 
-Note that given a generalized exact outcome, the exact outcome built from it will be equivalent.
-Importantly, this notion of equivalence is in fact an equivalence relation (i.e.
-a relation with reflexivity, symmetry, and transitivity):
-
-\begin{prop}
-The equivalence $\sim$ above is an equivalence relation.
-\end{prop}
+Of course, the nature of the set $X$ will vary greatly depending on the behavior of the collection $\mathcal{S}$ of observations.
+The following function will provide us with a collection of distinguished subsets of $X$.
+Define the \emph{verification function} $U:\mathcal{S}\to 2^X$, the set of subsets of $X$, by:
+$$
+U(\mathsf{o}) = \{s\in X\ |\ \mathsf{o}\text{ is successful if }s\text{ is true}\}.
+$$
+This function behaves nicely with our algebra of observations:
+\begin{lem}
+\label{ubehavior}
+Let $\mathsf{o}_1,\mathsf{o}_2,\ldots$ be some collection of observations. Then we have:
+$$U(\bigwedge\limits_{i=1}^{n} \mathsf{o}_i) = \bigcap\limits_{i=1}^{n} U (\mathsf{o}_i) \text{ and }U(\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i) = \bigcup\limits_{i=1}^{\infty} U (\mathsf{o}_i).$$
+That is, $U$ respects finite conjunction and countable disjunction.
+\end{lem}
 \begin{proof}
-Let $\mathcal{P},\mathcal{P}',\mathcal{P}''$ be exact outcomes. Reflexivity and symmetry are immediate since not being mutually exclusive is reflexive and symmetric.
-
-For transitivity, suppose $\mathcal{P}\sim\mathcal{P}'$ and $\mathcal{P}'\sim\mathcal{P}''$, and pick some positive integers $m$ and $n$.
-We will show that $\mathsf{o}_m$ is not mutually exclusive to $\mathsf{o}''_n$.
-By hypothesis, for every $\mathsf{o}'_j\in\mathcal{P}'$, we have $\mathsf{o}_m$ and $\mathsf{o}'_j$ are not mutually exclusive and $\mathsf{o}'_j$ and $\mathsf{o}''_n$ are not mutually exclusive.
-Because $\mathcal{P}'$ is an exact outcome and neither $\mathsf{o}_m$ nor $\mathsf{o}''_n$ are ever mutually exclusive from any $\mathsf{o}'_j$, we may select an integer $J$ large enough so that $\mathsf{o}'_J\implies\mathsf{o}_m$ and $\mathsf{o}'_J\implies\mathsf{o}''_n$.
-Hence whenever $\mathsf{o}'_J$ is verified, both $\mathsf{o}_m$ and $\mathsf{o}''_n$ are verified, so they are not mutually exclusive.
-Because this holds for all positive integers $m,n$, we have that $\mathcal{P}_1\sim\mathcal{P}_3$.
+Notice that $s\in U(\bigwedge\limits_{i=1}^{n} \mathsf{o}_i)$ if and only if all of the $\mathsf{o}_i$ are successful whenever $s$ is true.
+This is equivalent to the right-hand side, i.e.
+$s\in \bigcap\limits_{i=1}^{n} U (\mathsf{o}_i)$.
+The other equality is similarly straightforward.
 \end{proof}
 
-Next, we will define our universe of discourse.
+\begin{thm}
+For an experimental domain $\mathcal{S}$ with set of possibilities $X$ and verification function $U$, the collection $\mathsf{T} = U(\mathcal{S})\subset2^X$ is a Hausdorff, second-countable topology on $X$.
+\end{thm}
+\begin{proof}
+First, the null obervation will map to the empty set, and the trivial observation will map to all of $X$.
+Next, we must check arbitrary union and finite intersections.
+Recall that the experimental domain $\mathcal{S}$ has a countable basis $\mathcal{B} = \{\mathsf{o}_1,\mathsf{o}_2,\ldots\}$.
 
-\begin{defn}
-	A \textbf{set of possibilities} $X$ for an experimental domain $\mathcal{S}$ is the collection $\mathcal{C}$ of all exact outcomes for $\mathcal{S}$, modulo the equivalence relation $\sim$ defined above.
-That is, $X=\mathcal{C}/\sim$.
-\end{defn}
+Let $\{U_i\ |\ i\in I\}\subseteq \mathsf{T}$, where $I$ is some indexing set.
+Let $U_i = U(\mathsf{O}_i)$, where each $\mathsf{O}_i = \bigcup\limits_{j=1}^{\infty}\mathsf{o}_{n_{i,j}}$ for appropriately indexed basis elements $\mathsf{o}_{n_{i,j}}$, where $n_{i,j}$ is a positive integer.
+Now, since there are only countably many basis elements, we know that there are only countably many distinct $n_{i,j}$'s, so we may define a sequence of positive integers $m_k, k=1,2,\ldots$ such that $$\{m_k\}_{k=1}^{\infty} = \{n_{i,j}\ |\ i\in I, j=1,2,\ldots\}.$$ Then using Lemma \ref{ubehavior}, we have the following:
+$$
+\bigcup\limits_{i\in I}U(\mathsf{O}_i) = \bigcup\limits_{i\in I}\bigcup\limits_{j=1}^{\infty}U(\mathsf{o}_{n_{i,j}}) = \bigcup\limits_{k=1}^{\infty}U(\mathsf{o}_{m_k}) = U(\bigvee\limits_{k=1}^{\infty} \mathsf{o}_{m_k})
+$$
+which is an element of $\mathsf{T}$, so this collection is closed under arbitrary union.
 
-Equivalently, a set of possibilities $X$ for $\mathcal{S}$ may be specified by choosing one exact outcome from each equivalence class.
-We will now provide a notion of containment for exact outcomes being ``within" an experimental observation.
+The proof for finite intersection is simpler. Let $\{U_i\}_{i=1}^N$ be a finite collection of elements of $\mathsf{T}$. Then as above writing $U_i = U(\mathsf{O}_i)$, we have:
+$$
+\bigcap\limits_{i=1}^N U(\mathsf{O}_i) = U(\bigwedge\limits_{i=1}^{N} \mathsf{O}_i)
+$$
+by Lemma \ref{ubehavior}, as required. Hence $\mathsf{T}$ is a topology. In what follows we will refer to elements of $\mathsf{T}$ as open sets.
 
-\begin{defn}
-An exact outcome $\mathcal{P}=\{\mathsf{o}_i\}_{i=1}^{\infty}$ for experimental domain $\mathcal{S}$ is said to be an element of an experimental observation $\mathsf{o}\in\mathcal{S}$ if there exists some $N$ such that for all $n>N$, we have $\mathsf{o}_n\implies\mathsf{o}$.
-We denote this as $\mathcal{P}\in\mathsf{o}$.
-\end{defn}
+It is immediate that this topology is second-countable, since by Lemma \ref{ubehavior} the countable basis of observations for $\mathcal{S}$ becomes under $U$ a countable basis for the topology $\mathsf{T}$.
+To see why it is Hausdorff, let $x,y\in X$ be distinct possibilities.
+From Definition \ref{possibilitiesdef}, we have several cases to consider; however, it turns out that only one is possible.
+Without loss of generality, suppose $x\imp y$.
+But then we have $x\comp y$, so by symmetry we must also have $y\imp x$; in other words, $x=y$, contrary to assumption.
+It follows that $x\ncomp y$.
+Now suppose for a contradiction that for every open set $U\ni x$ and $V\ni y$ we have $U\cap V\neq \emptyset$.
+TODO: PROBLEM
 
-In light of the above definition, we will refer to experimental observations either as observations or as sets.
-
-\begin{prop}
-Let $\mathcal{P}=\{\mathsf{o}_i\}_{i=1}^{\infty}$ be an exact outcome which is an element of observation $\mathsf{o}$.
-Suppose $\mathcal{P}'$ is another exact outcome $\mathcal{P}'=\{\mathsf{o}'_j\}_{j=1}^{\infty}$ such that $\mathcal{P}\sim\mathcal{P}'$.
-Then $\mathcal{P}'\in\mathsf{o}$.
-\end{prop}
-\begin{proof}
-Suppose not - then we must be in case (1) of the definition of (generalized) exact outcome.
-Then there exists some $N_1$ such that for all $n>N_1$, we have $\mathsf{o}'_n\asymp\mathsf{o}$.
-By hypothesis, there exists some $N_2$ such that for all $m>N_2$, we have $\mathsf{o}_m\implies\mathsf{o}$.
-Let $N=\max(N_1,N_2)$.
-But then for all $k>N$, we have that $\mathsf{o}_k$ and $\mathsf{o}'_k$ are not mutually exclusive.
-This contradicts $\mathsf{o}'_k\asymp\mathsf{o}$.
 \end{proof}
-
-In particular, we immediately extend the definition of ``exact outcome containment" for elements of $\mathcal{S}$ to containment of equivalence classes of exact outcomes in $\mathcal{C}/\sim$.
-By the above, we can use containment of an individual exact outcome interchangeably with its equivalence class.
-In practice, most of our work will use individual exact outcomes representing an equivalence class, while the technical definitions will use their equivalence classes.
-
-\begin{prop}
-Let $\{\mathsf{o}_i\}_{i=1}^{n}$ be some collection of experimental observations.
-Then the conjunction $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis\bigwedge\limits_{i=1}^{n} \mathsf{s}_i, \mathsf{e}_{\wedge}(\mathsf{e}_i)\rrparenthesis$ contains precisely the exact outcomes that are in all of the $\mathsf{o}_i$'s.
-\end{prop}
-\begin{proof}
-Let $\mathcal{P} = \{\mathsf{o}'_i\}_{i=1}^{\infty}$ be an exact outcome, and suppose $\mathcal{P}\in\mathsf{o}_i$ for $i=1,\ldots,n$.
-Then there exists $N$ large enough so that $\mathsf{o}'_m\implies\mathsf{o}_i$ for all $m>N$ and all $i=1,\ldots,n$, which is possible since $n$ is finite.
-It follows that $\mathsf{o}'_m\implies\bigwedge\limits_{i=1}^{n} \mathsf{o}_i$ for all $m>N$, so $\mathcal{P}\in\bigwedge\limits_{i=1}^{n} \mathsf{o}_i$, as desired.
-
-Next, let $\mathcal{P}\in\bigwedge\limits_{i=1}^{n} \mathsf{o}_i$. It follows that $\mathcal{P}\in\mathsf{o}_i$ for each $i$ immediately from the definitions. 
-\end{proof}
-
-\begin{prop}
-Let $\{\mathsf{o}_i\}_{i=1}^{\infty}$ be some collection of experimental observations.
-Then the disjunction $\bigvee\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis\bigvee\limits_{i=1}^{n} \mathsf{s}_i, \mathsf{e}_{\vee}(\mathsf{e}_i)\rrparenthesis$ contains precisely the exact outcomes that are in any of the $\mathsf{o}_i$'s.
-\end{prop}
-\begin{proof}
-Let $\mathcal{P} = \{\mathsf{o}'_i\}_{i=1}^{\infty}$ be an exact outcome, and suppose $\mathcal{P}\in\mathsf{o}_k$ for some $k\geq1$.
-Then there exists $N$ large enough so that $\mathsf{o}'_m\implies\mathsf{o}_k$ for all $m>N$.
-It follows that $\mathsf{o}'_m\implies\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i$ for all $m>N$, so $\mathcal{P}\in\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i$, as desired.
-
-Next, let $\mathcal{P}\in\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i$.
-Then there exists $N$ such that for all $m>N$, we have $\mathsf{o}'_m\implies \bigvee\limits_{i=1}^{\infty} \mathsf{o}_i$.
-Pick some such $m$, and so in particular, there exists $k$ such that $\mathsf{o}'_m\implies\mathsf{o}_k$.
-By definition of (non-generalized) exact outcome, this means $\mathcal{P}\in \mathsf{o}_k$, as required.
-\end{proof}
-
-TODO: check this. the definition of experimental observation and their OR and AND are a little hard to use as needed here.
-
-TODO: maybe make a new notation for observation as a set of outcomes.
-
-TODO: a lot of the indices/notation might be a little hard to read/follow.
-
-We are now ready to prove one of the main results of this chapter.
-The following theorem links the philosophical definitions at the beginning of the chapter to the mathematical structure of a topological space.
-
-TODO: maybe add a theorem environment for this one
-
-\begin{prop}
-Let $\mathcal{S}$ be an experimental domain with (by definition countable) basis $\mathcal{B}$, and let $X = \mathcal{C}/\sim$ be the set of possibilities.
-Then $(X,\mathcal{S})$ is a Hausdorff, second-countable topological space with points $X$ and open sets $\mathcal{S}$, where union and intersection are given by conjunction and disjunction of experimental observations, respectively.
-\end{prop}
-\begin{proof}
-We first will show that the collection of sets $\mathsf{o}\in\mathcal{S}$ do indeed form a topology on $X$.
-First, the experimental observation formed by the (countable) disjunction of all observations in $\mathcal{B}$ is an observation which contains all others; that is, this set contains all of $X$.
-The empty observation contains no elements of $X$, so serves as the empty set.
-
-Next, let $\mathsf{o}_1,\mathsf{o}_2\in\mathcal{S}$.
-Then by the previous proposition about conjunction, we have that the set of exact outcomes corresponding to $\mathsf{o}_1\wedge\mathsf{o}_2$ is precisely $\mathsf{o}_1\cap\mathsf{o}_2$.
-For a countable subset of $\mathcal{S}$, the same argument shows that their union (as sets of exact outcomes) matches the disjunction (as experimental observations).
-For arbitrary collections, because $\mathcal{B}$ is countable and generates all of the experimental observations, we can rewrite an arbitrary union of the corresponding subsets of exact outcomes as a countable disjunction of experimental observations, and we are reduced to the countable case.
-
-This proves that $(X,\mathcal{S})$ is a topology.
-Further, we can see that $\mathcal{B}$ is a countable basis for the topology (because conjunction and disjunction correspond to intersection and union), so it is second-countable.
-Next, we prove that $(X,\mathcal{S})$ is Hausdorff.
-
-Let $\mathcal{P} = \{\mathsf{o}_i\}_{i=1}^{\infty}$ and $\mathcal{P}' = \{\mathsf{o}'_i\}_{j=1}^{\infty}$ be two non-equivalent exact outcomes.
-Then there exist some $m,n$ such that $\mathsf{o}_m\asymp\mathsf{o}'_n$.
-But then $\mathcal{P}\in\mathsf{o}_m$ and $\mathcal{P}'\in\mathsf{o}'_n$, which are two disjoint sets in $\mathcal{S}$.
-This completes the proof.
-\end{proof}
-
-\begin{defn}
-	A \textbf{verifiable set} $U \subseteq X$ is a subset of possibilities for which there exists an experimental observation $\mathsf{o}\in\mathcal{S}$  = (\text{``The object is in } U \text{"}, $\mathsf{e}_\in(U))$ where $\mathsf{e}_\in(U)$ is an experimental test that succeeds only if the object to identify is an element of $U$.
-\end{defn}
-
-\begin{defn}
-	A \textbf{refutable set} $U \subseteq X$ is a subset of possibilities for which there exists an experimental counter-observation $\mathsf{o} = (\text{``The object is in } U \text{"}, \mathsf{e}_{\notin}(U))$ where $\mathsf{e}_{\notin}(U)$ is an experimental test that succeeds only if the object to identify is not an element of $U$.
-\end{defn}
-
-\begin{prop}
-	The complement $U^C$ of a verifiable set $U \subseteq X$ is a refutable set.
-\end{prop}
 

### 2017-07-13 Articles/TopologyAndExperimentalDistinguishability/article.tex
Changed author ordering and corresponding author
Title, abstract, intro, and conclusion edits

--- 2017-07-12 Articles/TopologyAndExperimentalDistinguishability/article.tex
@@
 
-\title{Topology and physical distinguishability}
+\title{Topology and experimental distinguishability}
 
 %% Group authors per affiliation:
-\author[]{Gabriele Carcassi\corref{cor1}}
+\author[]{Christine A. Aidala\corref{cor1}}
+\ead{caidala@umich.edu}
+
+\author[]{Gabriele Carcassi}
 \ead{carcassi@umich.edu}
-
-\author[]{Christine A. Aidala}
-\ead{caidala@umich.edu}
 
@@
 \begin{abstract}
-We formalize the relationship between topological spaces and the ability to distinguish objects experimentally, providing understanding and justification as to why topological spaces and continuous functions are pervasive tools in the physical sciences.
-The aim is to use these ideas as a stepping stone to give a more rigorous physical foundation to dynamical systems and, in particular, Hamiltonian dynamics.
+We formalize the relationship between topological spaces and the ability to distinguish objects experimentally, providing understanding and justification as to why topological spaces and continuous functions are pervasive tools in science.
 We first define an experimental observation as a statement that can be verified using an experimental procedure and show that observations are closed under finite conjunction and countable disjunction.
@@
 \begin{keyword}
-topological spaces \sep other keyword
-\MSC[2010] 00-01\sep  99-00
+topological spaces \sep TODO
+\MSC[2010] TODO \sep TODO
 \end{keyword}
@@
 Why is it so successful?
-What physical property is captured by topological spaces that is so fundamental for scientific investigation?
+What property is captured by topological spaces that is so fundamental for scientific investigation?
 
 We believe we have identified the answer: topologies capture the way objects are experimentally distinguishable.
-It is open sets that are associated with experimental outcomes, not the values themselves.
+It is open sets that are associated with experimental outcomes, not the points themselves.
 For example, as real values are only measurable with finite precision, the standard topology on Euclidean spaces is the one that is physically meaningful.
@@
 
-We believe this work not only provides interesting insights, but a firm justification of the use of topological ideas in the physical sciences.
+We believe this work not only provides interesting insights, but a firm justification of the use of topological ideas in the sciences.
 
@@
 
-\section{Conclusion}
-
-We have shown that it is possible to formally capture the requirements of experimental sciences using experimental observations and their properties.
-These definitions lead naturally to topological spaces, giving the result that any set of experimentally distinguishable possibilities is a Hausdorff and second-countable topological space.
-In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects, as they too can be given a Hausdorff and secound-countable topology.
-The ability to create functions of any order confirms that the universe of discourse provided by these definitions is closed and self-consistent.
-
-We hope that this work, because of both its method and result, can provide a more solid foundation to formalize experimental sciences.
-
 \begin{table}[h]
@@
 
+\section{Conclusion}
+
+We have shown that it is possible to formally capture the requirements of experimental sciences using experimental observations and their properties.
+These definitions lead naturally to topological spaces, giving the result that any set of experimentally distinguishable possibilities is a Hausdorff and second-countable topological space.
+In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects, as they too can be given a Hausdorff and secound-countable topology.
+The ability to create functions of any order confirms that the universe of discourse provided by these definitions is closed and self-consistent.
+
+We hope that this work, because of both its method and result, can provide a more solid foundation to formalize experimental sciences.
+
 \section*{Acknowledgments}

### 2017-07-13 Articles/TopologyAndExperimentalDistinguishability/article.tex

--- 2017-07-13 Articles/TopologyAndExperimentalDistinguishability/article.tex
@@
 
-\title{Topology and physical distinguishability}
+\title{Topology and experimental distinguishability}
 
@@
 
-\author[umich]{Christine A. Aidala}
+\author[]{Christine A. Aidala\corref{cor1}}
 \ead{caidala@umich.edu}
 
-\author[umich]{Gabriele Carcassi\corref{cor1}}
+\author[]{Gabriele Carcassi}
 \ead{carcassi@umich.edu}
 
-\author[umich]{Mark J. Greenfield}
+\author[]{Mark J. Greenfield}
 \ead{markjg@umich.edu}
 
-\address[umich]{University of Michigan, Ann Arbor, MI 48109, USA}
+\address{University of Michigan, Ann Arbor, MI 48109, USA}
 
@@
 \begin{abstract}
-We formalize the relationship between topological spaces and the ability to distinguish objects experimentally, providing understanding and justification as to why topological spaces and continuous functions are pervasive tools in the physical sciences.
-The aim is to use these ideas as a stepping stone to give a more rigorous physical foundation to dynamical systems and, in particular, Hamiltonian dynamics.
-We first define an experimental observation as a statement that can be verified using an experimental procedure.
-We show that observations are not closed under negation and countable conjunction, but are closed under finite conjunction and countable disjunction.
-We then consider observations that identify elements in a set and show how they induce a Hausdorff and second countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
-We then show that only continuous functions preserve experimental distinguishability and that the collection of these functions can be given a Hausdorff and second countable topology.
-This shows that the universe of discourse of experimental distinguishability so defined is closed and consistent.
+We formalize the relationship between topological spaces and the ability to distinguish objects experimentally, providing understanding and justification as to why topological spaces and continuous functions are pervasive tools in science.
+We first define an experimental observation as a statement that can be verified using an experimental procedure and show that observations are closed under finite conjunction and countable disjunction.
+We then consider observations that identify elements in a set and show how they induce a Hausdorff and second-countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
+We then show that only continuous functions preserve experimental distinguishability and that the collection of these functions can be given a Hausdorff and second-countable topology.
 
@@
 Why is it so successful?
-What physical property is captured by topological spaces that is so fundamental for scientific investigation?
+What property is captured by topological spaces that is so fundamental for scientific investigation?
 
 We believe we have identified the answer: topologies capture the way objects are experimentally distinguishable.
-It is open sets that are associated with experimental outcomes, not the values themselves.
-For example, as real values are only measurable with finite precisions, the standard topology on euclidean spaces is the one that is physically meaningful.
-Along the same line, experimental relationships will need to preserve what can be experimentally distinguished and they will therefore be continuous functions.
+It is open sets that are associated with experimental outcomes, not the points themselves.
+For example, as real values are only measurable with finite precision, the standard topology on Euclidean spaces is the one that is physically meaningful.
+Along the same line, experimental relationships need to preserve what can be experimentally distinguished, and they are therefore represented as continuous functions.
 
-In this sense, experimentally distinguishability is \emph{the} application of topology in experimental science.
-It is fitting, then, that topology is so wide-spread as defining what can be experimentally identified is truly a fundamental aspect in science.
-The aim of this paper is to lay down a formal framework that formalizes this insight.
+In this sense, experimental distinguishability is \emph{the} application of topology in experimental science.
+It is fitting, then, that topology is so widespread given that defining what can be experimentally identified is truly a fundamental aspect of science.
+The aim of this paper is to lay down a framework that formalizes this insight.
 
 The work is organized as follows.
-We will first define experimental observations as statements paired with an experimental test that is able to verify them.
-We will study their properties under logical operations and show they are only closed under finite conjunction and countable disjunction.
+We first define experimental observations as statements paired with an experimental test that is able to verify them.
+We study their properties under logical operations and show that they are only closed under finite conjunction and countable disjunction.
+We then show that any set of objects that are experimentally distinguishable is a Hausdorff second-countable topological space.
+Both the points and the topology of the space are formally defined in terms of the observations themselves.
+Finally, we study relationships between experimentally distinguishable objects.
+These are represented by continuous functions, and they can be shown to be themselves physically distinguishable.
 
-We will then show that any set of objects that are experimentally distinguishable is a Hausdorff second-countable topological space.
-Both the points and the topology of the space will be formally defined in terms of the observations themselves.
-
-Finally, we will study relationships between experimentally distinguishable objects. This will be represented by continuous functions and can be shown to be themselves physically distinguishable
-
-We believe this work not only provides interest insights, but a firm justification of the use of topological ideas in the physical science.
+We believe this work not only provides interesting insights, but a firm justification of the use of topological ideas in the sciences.
 
@@
 
-\section{Conclusion}
-
-We have shown that it is possible to formally capture the requirements of experimental sciences using experimental observations and their properties.
-These definitions lead naturally to topological spaces giving the result that any set of experimentally distinguishable possibilities is a Hausdorf and second-countable topological space.
-In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects as they too can be given a Hausdorf and secound-countable topology.
-The ability to create functions of any order confirms that the universe of discourse provided by these definition is closed and self-consistent.
-
-We hope that this work, because of both its method and result, can provide a more solid foundation to formalize experimental sciences.
-
 \begin{table}[h]

### 2017-07-14 Blog/07-WhatIsInformationEntropy.tex
Blog post review

@@
 it measures "disorder", it has to do with how much the system is homogeneous, ...).
-Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear some defining it as "lack of information" or any other vague description.
+Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear people describing it as "lack of information" or any other vague description.
 So, let's see if we can do better.
@@
 
-Suppose you have a bag with 40 balls.
-20 are green and 20 are red.
+Suppose you have a bag with forty balls.
+Twenty are green and twenty are red.
 Suppose you pick a ball and I want to know the color.
@@
 The answer is one bit: 0 for green and 1 for red.
-That is: all the information can be encoded into one zero-one choice.
+That is: all the information can be encoded into a single zero-one choice.
 
-Let's do it again in another case. Suppose all 40 balls are green. How many bits of information? The answer is zero bits: I already know you are going to get a green ball.
+Let's do it again in another case. Suppose all forty balls are green. How many bits of information? The answer is zero bits: I already know you are going to get a green ball.
 
-Suppose there are 10 balls for each of these color: green, red, yellow and blue.
+Suppose there are ten balls for each of these color: green, red, yellow and blue.
 How many bits of information?
 We can set 00 for green, 01 for red, 10 for yellow and 11 for blue.
-That is two zero-one choices: it's two bits of information.
+Four cases can be encoded into two zero-one choices: it's two bits of information.
 
@@
 So, in this case more entropy correspond to less knowledge.
-But this idea does no generalize.
+But this idea does not generalize.
 
@@
 
-This hopefully helps you understand what information entropy measures: how many bits of informations are required on average to identify an element in a distribution.
+At the end of the day, information entropy is a number: to understand it you need to understand what that number quantifies and in what units.
+If do not have a clear idea what that number is measuring, you do not understand what it is.
+That number measures how many bits of informations are required on average to identify an element in a distribution.
 Not "knowledge" or "lack of information" or "disorder" or any other vague concept.
-In the end, information entropy is a number: to understand it you need to understand what that number quantifies and in what units.
 

### 2017-07-14 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Skeleton for hamiltonian entropy paper

\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{dutchcal}
\usepackage{braket}
\usepackage{enumitem}

\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{calculator}
\usepackage{standalone}

\begin{document}

\title{DRAFT \\ Hamiltonian mechanics is conservation of information \\ (or why isolated systems conserve energy)}
\author{Christine A. Aidala, Gabriele Carcassi}

%\ifjournal
%	% As the page margins are so large, the full title does not fit
%	\titlerunning{From physical principles to classical and quantum particle mechanics}
%\fi

\date{\today}

\begin{abstract}
In this work we show that the canonical transformations of classical Hamiltonian mechanics are exactly the transformations that preserve information.
\end{abstract}

\maketitle

\section{Introduction}

\section{Single degree of freedom}
Define distribution over phase space.

Show change under coordinate transformation.

Define information entropy.

Show change under coordinate transformation.

Show that conservation of information entropy is equivalent to transformation having Jacobian equals to 1.

Show that canonical transformations are the ones and only the ones that have the Jacobian equals to one. Show relation to Poisson bracket. Show relation to incompressible flow. Show relation to symplectic form.

\section{Multiple degree of freedom}

Show conceptually what needs to happen. Need to be able to define marginal distributions on each degree of freedom. Jacobian preserved under all degrees of freedom. Symplectic form in multiple dimension preserves area and orthogonality. Product of marginals needs to be preserved.

Show that setting up the problem mathematically is more problematic because: symplectic geometry does not have the tools to talk about distributions as it focuses on objects that are coordinate independent, statistics does not care that distribution and information entropy is coordinate dependent since the prime object is the cumulative distribution function.

Spend the rest at making links between geometric objects and distributions.

Show the definitions for mutual information entropy.

\section{Discussion}

Why should we care? Because it provides links.

Uncertainty principle as relationship between entropy and uncertainty.

Relativistic Hamiltonian motion as preservation of entropy.

Relationship to conservation of energy in isolated systems.

\begin{thebibliography}{0}

\end{thebibliography}

\end{document}

### 2017-07-15 Blog/07-WhatIsInformationEntropy.tex
Final edit to entropy blog post

@@
 
-TL;DR Information entropy tells you the average number of bits (i.e. answers to yes/no questions) you need to identify an element from a distribution.
+TL;DR Information entropy tells you the average number of bits (i.e.~answers to yes/no questions) you need to identify an element from a distribution.
 
@@
 The concept of entropy in physics is unfortunately associated with a certain amount of vagueness.
-While mathematically is well defined, I have never heard a crisp conceptual understanding (e.g.
-it measures "disorder", it has to do with how much the system is homogeneous, ...).
-Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear people describing it as "lack of information" or any other vague description.
+While mathematically it is well defined, I have never heard a crisp conceptual understanding (e.g.
+it measures ``disorder", it has to do with how much the system is homogeneous, something to do with the number of microstates, ...).
+Entropy in information theory, or Shannon's entropy, instead has a crisp intuitive meaning and therefore I cringe when I hear people describing it as ``lack of information" or any other vague description.
 So, let's see if we can do better.
@@
 You pick an element according to that distribution.
-On average, how many bits of information you have to use to tell me which element have you picked?
+On average, how many bits of information do you have to use to tell me which element you have picked?
 How many one-zeros?
@@
 How many questions do I have to ask you?
-One suffices: "is it red?".
-If the answer is yes, I know it's red, if the answer is no, I know it's green.
+One suffices: ``Is it red?" If the answer is yes, I know it's red, if the answer is no, I know it's green.
 We can rephrase by asking: how many one-zero choices?
@@
 
-Suppose there are ten balls for each of these color: green, red, yellow and blue.
+Suppose there are ten balls for each of these colors: green, red, yellow and blue.
 How many bits of information?
@@
 
-Shannon's entropy simply generalizes this idea: given a $n$ cases each with a probability $p_i$, the average number of bits (i.e. yes/no answers) required to identify an element is:
+Shannon's entropy simply generalizes this idea: given $n$ cases each with a probability $p_i$, the average number of bits (i.e. yes/no answers) required to identify an element is:
 
@@
 If it's base $2$, then we are using bits.
-But sometimes is useful to use base $e$ (i.e.
-the natural log) since it has nice properties.
-In that case we are using nats, and we have $1$ bit = $\ln 2$ nats.
+But sometimes it is useful to use base $e$ (i.e.~the natural log) since it has nice properties.
+In that case we are using nats (as in natural log), and we have $1$ bit = $\ln 2$ nats.
 
@@
 	\item $H(p_1, ..., p_n)$ is continuous.
-	\item $H(p_1, ..., p_n)$ increases if the number of choice increases.
+	\item $H(p_1, ..., p_n)$ increases if the number of choices increases.
 That is, suppose you have two uniform distributions over $n$ and $m$ cases respectively.
@@
 
-While we are not going to derive it here, the derivation is not particularly insightful, we can still understand why it is what it is.
-If a particular case is a choice among $n$ equal cases, it's probability is $p_i=\frac{1}{n}$ and the number of yes/no questions we have to make to get to it is $\log_2 n = \log_2 \frac{1}{p_i} = - \log_2 p_i$.
+Since the derivation is not particularly insightful, we are not going to derive it here.
+But we can still understand why it is what it is.
+If a particular case is a choice among $n$ equal cases, its probability is $p_i=\frac{1}{n}$ and the number of yes/no questions we have to make to get to it is $\log_2 n = \log_2 \frac{1}{p_i} = - \log_2 p_i$.
 So, when that particular case is selected, we need $- \log_2 p_i$ bits.
-Now we take the expectation, the average over all cases, and we have $H(p_1, ..., p_n) = - \sum_{i=1}^{n} p_i \log p_i$.
-We now have the average value of yes/no questions (i.e.
-bits) needed to identify an element in the distribution.
+Now we take the average over all cases and we have $H(p_1, ..., p_n) = \sum_{i=1}^{n} p_i \, (- \log_2 p_i)$.
+We now have the average number of yes/no questions (i.e.~bits) needed to identify an element in the distribution.
 
@@
 
-There are a couple of aspects we need to clarify as it allows to understand why such concept is useful in physics.
+There are a couple of aspects we need to clarify to understand why this concept is useful in physics.
 
-First of all, information entropy has nothing to do with knowledge (e.g.
-what we know about the system or what can be known).
-This misunderstanding arises because the concept can be applied to probability distribution that represent our knowledge of the system, and in some cases it may relate to knowledge.
+First of all, information entropy has nothing to do with knowledge (e.g.~what we know about the system or what can be known).
+This misunderstanding arises because the concept can be applied to probability distributions that represent our knowledge of the system, and in some cases it may seem to relate to knowledge.
 For example, if there are four boxes and I do not know in which one the ball is hiding, I have a uniform distribution on four cases which has 2 bits of information entropy.
 Conversely, if I know where it is, the entropy is 0 bits.
-So, in this case more entropy correspond to less knowledge.
-But this idea does not generalize.
+So, in this case more entropy corresponds to less knowledge.
+But this is purely coincidental.
 
@@
 For example, we could have a mass distribution in space.
-In this case, information entropy does not equate to lack of knowledge.
+In this case, information entropy clearly cannot correspond to lack of knowledge.
 Our knowledge of the mass distribution is perfect: we know exactly how it is spread over a region.
-The information entropy of a mass distribution in space is as much a physical property of the system as it is the mass distribution itself and it tells us how many bits are necessary to identify the position of a part of the system.
+The information entropy of a mass distribution in space is as much a physical property of the system as is the mass distribution itself and it tells us how many bits are necessary to identify the position of a part of the system.
 Unlike the previous case, this is an objective number.
@@
 At the end of the day, information entropy is a number: to understand it you need to understand what that number quantifies and in what units.
-If do not have a clear idea what that number is measuring, you do not understand what it is.
+If you do not have a clear idea what that number is measuring, you do not understand what it is.
 That number measures how many bits of informations are required on average to identify an element in a distribution.
-Not "knowledge" or "lack of information" or "disorder" or any other vague concept.
+Not ``knowledge" or ``lack of information" or ``disorder" or any other vague concept.
 

### 2017-07-16 Blog/08-HamiltonianMechanicsIsConservationOfInformationEntropy.tex
Reviewed post

@@
 
-This is one of those result that I figured out a few years ago but I still don't know if and where to publish it.
-Fortunately for me, I care a lot more about figuring things out than publishing papers.
-
 The idea is the following: suppose you have a distribution over position and momentum $\rho(x, p)$.
 Suppose you evolve it in time and get a final distribution $\hat{\rho}(x, p)$ using Hamiltonian evolution.
-That is: you take each little volume and move it according to the trajectories given by Hamiltonian mechanics.
+That is: you take each little area in phase space and move it according to Hamilton's equations.
 How does the information entropy change?
@@
 That is: Hamiltonian systems are the ones and only the ones that preserve information entropy of a distribution $\rho(x, p)$ over states.
-That is: identifying an element in the initial distribution requires exactly the same information as an element in the final distribution.
-That is: knowing the initial element you know the final element and vice-versa.
-That is: the evolution is deterministic and reversible.
+Identifying an element in the initial distribution requires exactly the same information as identifying an element in the final distribution.
+Knowing the initial element you know the final element and vice-versa.
+The evolution is deterministic and reversible.
 
@@
 
-Because the evolution is Hamiltonian, the Liouville's theorem applies so the Jacobian is unitary. Let's verify that:
+Because the evolution is Hamiltonian, Liouville's theorem applies so the Jacobian is unitary.
+Let's verify that by combining the Jacobian definition in \eqref{newDistribution} with the new coordinates defined in \eqref{newCoordinates}:
 \begin{equation}
@@
 
-This means the volume $d\hat{x}\,d\hat{p} = dx \,dp$ remains unchanged and the final density at the final state is equal to the initial density at the initial state: $\rho(\hat{x}, \hat{p}) = \rho(x, p)$.
+This means the volume $d\hat{x}\,d\hat{p} = dx \,dp$ remains unchanged and the final density at the final state is equal to the initial density at the initial state: $\hat{\rho}(\hat{x}, \hat{p}) = \rho(x, p)$.
 
@@
 \begin{aligned}
--\int \hat{\rho} \log \hat{\rho} \; d\hat{x}\,d\hat{p}&= -\int \frac{\rho}{|J|} \log \frac{\rho}{|J|} \; |J| dx \,dp = -\int \frac{\rho}{|J|} \log \frac{\rho}{|J|} \; |J| dx \,dp \\
+-\int \hat{\rho} \log \hat{\rho} \; d\hat{x}\,d\hat{p}&= -\int \frac{\rho}{|J|} \log \frac{\rho}{|J|} \; |J| dx \,dp \\
 &= -\int \rho \log \frac{\rho}{|J|} \; dx \,dp \\
@@
 \end{equation}
-That is: the information entropy of the final distribution is equal to the one of the initial distribution plus the expectation of the Jacobian.
+That is: the information entropy of the final distribution is equal to the one of the initial distribution plus the expectation of the logarithm of the Jacobian.
 
-But since the evolution is Hamiltonian, the Jacobian is unitary. The information entropy of the final distribution is that of the initial distribution. The information entropy is conserved.
+But since the evolution is Hamiltonian, the Jacobian is unitary and its logarithm is therefore zero.
+The information entropy of the final distribution is that of the initial distribution.
+The information entropy is conserved.
 
@@
 We can run the argument backwards.
-If information entropy is conserved, the logarithm of the Jacobian has to be zero everywhere, which can only happen if the Jacobian is equal to one everywhere.
+Suppose we have a distribution over position and momentum $\rho(x, p)$.
+Suppose the evolution is such that the information entropy is conserved.
+How can we characterize the evolution?
+For the information entropy to be conserved the logarithm of the Jacobian has to be zero everywhere, which can only happen if the Jacobian is equal to one everywhere.
 Looking at \eqref{Jacobian}, this happens only if:
@@
 \dfrac{\partial}{\partial x} \left( \dfrac{dx}{dt} \right) + \dfrac{\partial}{\partial p} \left( \dfrac{dp}{dt} \right) = 0
-\label{Jacobian}
+\label{divergenceFree}
 \end{equation}
 
-This means that the vector field formed by the displacement vector $\left[ \dfrac{dx}{dt}, \dfrac{dp}{dt}\right]$ is divergence free. Which means it admits a potential $H$ such that:
+This means that the vector field formed by the displacement vector $\left[ \dfrac{dx}{dt}, \dfrac{dp}{dt}\right]$ is divergence free, which means it admits a potential $H$ such that:
 \begin{equation}
@@
 \section{Conclusion}
-As we have saw, Hamiltonian mechanics is equivalent to conservation of information entropy.
-This idea can be generalized to multiple independent degrees of freedoms.
-The math is a bit more involved but does not add to the general idea.
+As we have seen, Hamiltonian mechanics is equivalent to conservation of information entropy.
+This idea can be generalized to multiple independent degrees of freedom.
+The math is a bit more involved but the single degree of freedom already gives the general idea.
 

### 2017-07-18 Articles/TopologyAndExperimentalDistinguishability/article.tex
wrote out some of the old definitions in a more mathematical way

@@
 \begin{defn}
-	A \emph{statement} $\mathsf{s}$ is a declarative sentence that is either true or false, as in classical logic. Here, we allow for infinite statements. 
-\end{defn}
-
-In addition to the usual logical operations on statements, we introduce the following notation:
-
-\begin{defn}
-We say two statements $s$ and $t$ are \emph{compatible}, denoted $s\comp t$, if the conjunction $s\wedge t$ is not a contradiction; otherwise they are \emph{incompatible}, denoted $s\ncomp t$. 
+	A \textbf{statement} $\mathsf{s}$ is a declarative sentence that is either true or false, as in classical logic. 
 \end{defn}
@@
 \begin{defn}
-	An \emph{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e. it can be restarted and stopped an arbitrary number of times) that, if successful, terminates in a finite amount of time.
+	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e. it can be restarted and stopped an arbitrary number of times) that, if successful, terminates in a finite amount of time.
 \end{defn}
@@
 \begin{defn}
-	An \emph{experimental observation} is a tuple $\mathsf{o} = \llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is always successful.
-The experimental observation is \emph{verified} if the statement is true.
+	An \textbf{experimental observation} is a tuple $\mathsf{o} = \llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is always successful.
+The experimental observation is \textbf{verified} if the statement is true.
 \end{defn}
@@
 \begin{defn}
-	An \emph{experimental counter-observation} is a tuple $\mathsf{o}^c= \llbracket \mathsf{s}, \mathsf{e} \rrbracket$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is false if and only if the experimental test is always successful.
-The experimental counter-observation is \emph{refuted} if the statement is false.
+	An \textbf{experimental counter-observation} is a tuple $\mathsf{o}^c= \llbracket \mathsf{s}, \mathsf{e} \rrbracket$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is false if and only if the experimental test is always successful.
+The experimental counter-observation is \textbf{refuted} if the statement is false.
 \end{defn}
@@
 	\begin{itemize}
-	\item \emph{implication}: $\mathsf{o}_1 \imp \mathsf{o}_2$ ($\mathsf{o}_1$ implies $\mathsf{o}_2$) if $\mathsf{o}_2$ is always verified whenever $\mathsf{o}_1$ is verified
-	\item \emph{equivalence}: $\mathsf{o}_1 = \mathsf{o}_2$ ($\mathsf{o}_1$ is equal to $\mathsf{o}_2$) if $\mathsf{o}_1 \imp \mathsf{o}_2$ and $\mathsf{o}_2 \imp \mathsf{o}_1$
-	\item \emph{compatibility}: $\mathsf{o}_1 \comp \mathsf{o}_2$ ($\mathsf{o}_1$ is compatible with $\mathsf{o}_2$) if there exists a case in which $\mathsf{o}_1$ and $\mathsf{o}_2$ are jointly verified
+	\item \textbf{implication}: $\mathsf{o}_1 \imp \mathsf{o}_2$ ($\mathsf{o}_1$ implies $\mathsf{o}_2$) if $\mathsf{o}_2$ is always verified whenever $\mathsf{o}_1$ is verified
+	\item \textbf{equivalence}: $\mathsf{o}_1 = \mathsf{o}_2$ ($\mathsf{o}_1$ is equal to $\mathsf{o}_2$) if $\mathsf{o}_1 \imp \mathsf{o}_2$ and $\mathsf{o}_2 \imp \mathsf{o}_1$
+	\item \textbf{compatibility}: $\mathsf{o}_1 \comp \mathsf{o}_2$ ($\mathsf{o}_1$ is compatible with $\mathsf{o}_2$) if there exists a case in which $\mathsf{o}_1$ and $\mathsf{o}_2$ are jointly verified
 	\end{itemize}
@@
 \begin{defn}
-	The \emph{null observation} is any experimental observation that is never verified (i.e. the experimental test will never terminate successfully). 
+The \textbf{null observation}, denoted $\mathsf{o}_N$, is any experimental observation that is never verified (i.e. the experimental test will never terminate successfully).
 \end{defn}
@@
 \begin{defn}
-	The \emph{negation} or \emph{logical NOT} of an experimental observation $\mathsf{o}=\llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ is the experimental counter-observation $\neg \mathsf{o}=\llbracket\neg \mathsf{s}, \mathsf{e}\rrbracket$ where the negation of the statement $\mathsf{s}$ can be refuted by the experimental test $\mathsf{e}$.
+	The \textbf{negation} or \textbf{logical NOT} of an experimental observation $\mathsf{o}=\llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ is the experimental counter-observation $\neg \mathsf{o}=\llbracket\neg \mathsf{s}, \mathsf{e}\rrbracket$ where the negation of the statement $\mathsf{s}$ can be refuted by the experimental test $\mathsf{e}$.
 \end{defn}
@@
 \begin{defn}
-	The \emph{conjunction} or \emph{logical AND} of a finite collection  of experimental observations $\{\mathsf{o}_i\}_{i=1}^{n}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{n}$ is the experimental observation $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigwedge\limits_{i=1}^{n} \mathsf{s}_i$ is the conjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ is the experimental test that successfully terminates if and only if all $\{\mathsf{e}_i\}_{i=1}^{n}$ successfully terminate.
+	The \textbf{conjunction} or \textbf{logical AND} of a finite collection  of experimental observations $\{\mathsf{o}_i\}_{i=1}^{n}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{n}$ is the experimental observation $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigwedge\limits_{i=1}^{n} \mathsf{s}_i$ is the conjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ is the experimental test that successfully terminates if and only if all $\{\mathsf{e}_i\}_{i=1}^{n}$ successfully terminate.
 \end{defn}
@@
 \begin{defn}
-	The \emph{disjunction} or \emph{logical OR} of a countable (finite or infinite) collection of experimental observations $\{\mathsf{o}_i\}_{i=1}^{\infty}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{\infty}$ is the experimental observation $\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigvee\limits_{i=1}^{\infty} \mathsf{s}_i$ is the disjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ is the experimental test that successfully terminates if and only if at least one experimental test in $\{\mathsf{e}_i\}_{i=1}^{\infty}$ successfully terminates.
+	The \textbf{disjunction} or \textbf{logical OR} of a countable (finite or infinite) collection of experimental observations $\{\mathsf{o}_i\}_{i=1}^{\infty}=\{\llparenthesis \mathsf{s}_i, \mathsf{e}_i\rrparenthesis\}_{i=1}^{\infty}$ is the experimental observation $\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = \llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ where $\mathsf{s} = \bigvee\limits_{i=1}^{\infty} \mathsf{s}_i$ is the disjunction of the respective statements and $\mathsf{e} = \mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ is the experimental test that successfully terminates if and only if at least one experimental test in $\{\mathsf{e}_i\}_{i=1}^{\infty}$ successfully terminates.
 \end{defn}
@@
 
+\begin{rem}
+A useful equivalent characterization of $\mathsf{o}\imp\mathsf{o}'$ for observations $\mathsf{o}$ and $\mathsf{o}'$ is provided by the above constructions.
+We note that $\mathsf{o}\imp\mathsf{o}'$ if and only if $\mathsf{o}\wedge\mathsf{o}' = \mathsf{o}'$ if and only if $\mathsf{o}\vee\mathsf{o}'=\mathsf{o}$.
+\end{rem}
+
 \section{Experimental domain}
@@
 \begin{defn}
-	An \emph{experimental domain} is a set of observations closed under finite conjunction and countable disjunction such that all observations can be tested in infinite time. 
+	An \textbf{experimental domain} is a set of observations closed under finite conjunction and countable disjunction such that all observations can be tested in infinite time. 
 \end{defn}
@@
 \begin{defn}
-	A \emph{sub-basis} of an experimental domain is any subset that can generate all others via finite conjunction and countable disjunction.
-A \emph{basis} of an experimental domain is any subset that can generate all others by countable disjunctions.
+	A \textbf{sub-basis} of an experimental domain is any subset that can generate all others via finite conjunction and countable disjunction.
+A \textbf{basis} of an experimental domain is any subset that can generate all others by countable disjunctions.
 \end{defn}
@@
 \begin{rem}
-It is not hard to see that the (countable) disjunction of an entire basis produces an experimental observation which is always verified. Henceforth, we call this the \emph{trivial observation}. 
+It is not hard to see that the (countable) disjunction of an entire basis produces an experimental observation which is always verified.
+Henceforth, we call this the \textbf{trivial observation}, denoted $\mathsf{o}_T$.
 \end{rem}
@@
 
-We will now use our notions of experimental observations and experimental domains to construct a new space with a topology which will more rigorously capture our notion of ``exact" possibilities.
-We will begin with a given experimental domain $\mathcal{S}$.
-First, we will construct an intermediate space which will mimic the Borel algebra of a topological space:
+\begin{defn}
+A \textbf{set of possibilities} $(X,x^*)$ is a set together with a distinguished element $x^*$. 
+\end{defn}
 
 \begin{defn}
-Given an experimental domain $\mathcal{S}$, define the associated \emph{theoretical domain} $\mathcal{S}^*$ to be the closure under countable conjunction, disjunction, and negation of the set of statements in all experimental observations of $\mathcal{S}$.
+	A \textbf{verifiable set} $U \subseteq X$ is a subset of possibilities for which there exists an experimental observation $\mathsf{o} = (x^*\in U, \mathsf{e}_\in(U))$ where $\mathsf{e}_\in(U)$ is an experimental test that succeeds only if the object to identify is an element of $U$.
 \end{defn}
 
-Next, we will extract the appropriate subcollection of the theoretical domain in order to obtain what will serve as ``points" in our space:
+\begin{defn}
+	A \textbf{refutable set} $U \subseteq X$ is a subset of possibilities for which there exists an experimental counter-observation $\mathsf{o} = (x^*\in U, \mathsf{e}_{\notin}(U))$ where $\mathsf{e}_{\notin}(U)$ is an experimental test that succeeds only if the object to identify is not an element of $U$.
+\end{defn}
+
+\begin{lem}
+\label{complements}
+	The complement $U^C$ of a verifiable set $U \subseteq X$ is a refutable set.
+\end{lem}
+
+\begin{proof}
+	Let $U\subset X$ be a verifiable set.
+There exists an experimental observation $\mathsf{o} = (x^*\in U, \mathsf{e}_\in(U))$.
+Consider $\neg \mathsf{o} = (\neg x^*\in U, \mathsf{e}_\in(U)) = (x^*\notin U, \mathsf{e}_{\notin}(U^C)) = (x^*\in U^C, \mathsf{e}_{\notin}(U^C))$.
+That is, there exists a counter-observation $(x^*\in U^C, \mathsf{e}_{\notin}(U^C))$.
+Therefore $U^C$ is a refutable set.
+\end{proof}
+
+\begin{lem}
+\label{setbehavior}
+	Let $U_1, U_2, ...
+, U_n, ...$ be a countable infinite sequence of verifiable sets.
+The finite intersection $\bigcap\limits_{i=1}^{n} U_i$ and the countable union $\bigcup\limits_{i=1}^{\infty} U_i$ are verifiable sets.
+\end{lem}
+
+\begin{proof}
+	We claim the finite intersection of verifiable sets is a verifiable set.
+Let $U_1, U_2, ...
+, U_n \subseteq X$ be n verifiable sets.
+For each $U_i$ there exists an experimental observation $\mathsf{o}_i = (x^*\in U_i, \mathsf{e}_\in(U_i))$.
+Consider $\mathsf{o} = \bigwedge\limits_{i=1}^{n} \mathsf{o}_i = (\bigwedge\limits_{i=1}^{n} x^*\in U_i , \mathsf{e}_{\wedge}(\mathsf{e}_\in(U_i)))=( x^*\in \bigcap\limits_{i=1}^{n} U_i, \mathsf{e}_\in(\bigcap\limits_{i=1}^{n} U_i))$ is an experimental observation.
+We conclude $\bigcap\limits_{i=1}^{n} U_i$ is a verifiable set.
+
+	We claim the countable union of verifiable sets is a verifiable set.
+Let $U_1, U_2, ...
+, U_n, ...
+\subseteq X$ be an infinite sequence of verifiable sets.
+For each $U_i$ there exists an experimental observation $\mathsf{o}_i = (x^*\in U_i, \mathsf{e}_\in(U_i))$.
+Consider $\mathsf{o} = \bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = (\bigvee\limits_{i=1}^{\infty} x^*\in U_i, \mathsf{e}_{\vee}(\mathsf{e}_\in(U_i)))=( x^*\in \bigcup\limits_{i=1}^{\infty} U_i, \mathsf{e}_\in(\bigcup\limits_{i=1}^{\infty} U_i))$ is an experimental observation.
+Thus $\bigcup\limits_{i=1}^{\infty} U_i$ is a verifiable set.
+\end{proof}
+
+Next, we wish to isolate the possibility spaces for which there exists an appropriate collection of experimental observations.
+Let $(X,x^*)$ be a set of experimental possibilities.
+An \textbf{experimental identification} is any experimental observation of the form $(x^*\in S,\mathsf{e}_{\in}(S))$, where $S$ is any subset of $X$.
 
 \begin{defn}
-\label{possibilitiesdef}
-Given a theoretical domain $\mathcal{S}^*$, define the set $X$ of \emph{possibilities} to be the following subset:
-$$
-X = \{s\in\mathcal{S}^*\ |\ \forall t\in\mathcal{S}^*, \text{ either } s\imp t \text{ or } s\ncomp t\}
-$$
+	A set of experimental possibilities $X$ is \textbf{experimentally distinguishable} if there exists a collection $\mathsf{T}$ of experimental identifications which form an experimental domain such that given any two possibilities $x_1, x_2 \in X$ we can always find two mutually exclusive experimental observation such that each possibility is compatible with one observation.
 \end{defn}
 
-Of course, the nature of the set $X$ will vary greatly depending on the behavior of the collection $\mathcal{S}$ of observations.
-The following function will provide us with a collection of distinguished subsets of $X$.
-Define the \emph{verification function} $U:\mathcal{S}\to 2^X$, the set of subsets of $X$, by:
-$$
-U(\mathsf{o}) = \{s\in X\ |\ \mathsf{o}\text{ is successful if }s\text{ is true}\}.
-$$
-This function behaves nicely with our algebra of observations:
-\begin{lem}
-\label{ubehavior}
-Let $\mathsf{o}_1,\mathsf{o}_2,\ldots$ be some collection of observations. Then we have:
-$$U(\bigwedge\limits_{i=1}^{n} \mathsf{o}_i) = \bigcap\limits_{i=1}^{n} U (\mathsf{o}_i) \text{ and }U(\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i) = \bigcup\limits_{i=1}^{\infty} U (\mathsf{o}_i).$$
-That is, $U$ respects finite conjunction and countable disjunction.
-\end{lem}
+From the work above, we will quickly arrive at the following.
+
+\begin{thm}
+An experimentally distinguishable set $X$ together with its collection $\mathsf{T}$ of experimental identifications (realized as subsets of $X$) forms a Hausdorff, second-countable topology with open sets given by the verifiable sets.
+\end{thm}
 \begin{proof}
-Notice that $s\in U(\bigwedge\limits_{i=1}^{n} \mathsf{o}_i)$ if and only if all of the $\mathsf{o}_i$ are successful whenever $s$ is true.
-This is equivalent to the right-hand side, i.e.
-$s\in \bigcap\limits_{i=1}^{n} U (\mathsf{o}_i)$.
-The other equality is similarly straightforward.
+It is immediate that the trivial and null observations may be written as experimental identifications, and that they give $X\in\mathsf{T}$ and $\emptyset\in\mathsf{T}$.
+Proposition \ref{setbehavior} shows that the collection $\mathsf{T}$ is closed under finite intersection and countable union.
+Because $\mathsf{T}$ is given by an experimental domain of observations, there is a countable basis, so it is second-countable.
+To show it is closed under arbitrary union, notice that an arbitrary union may be rewritten as a union of basis elements, which is then a countable union, and so it remains in $\mathsf{T}$.
 \end{proof}
 
-\begin{thm}
-For an experimental domain $\mathcal{S}$ with set of possibilities $X$ and verification function $U$, the collection $\mathsf{T} = U(\mathcal{S})\subset2^X$ is a Hausdorff, second-countable topology on $X$.
-\end{thm}
+\begin{cor}
+The collection of refutable sets for $X$ is precisely the closed sets for the topology $\mathsf{T}$.
+\end{cor}
 \begin{proof}
-First, the null obervation will map to the empty set, and the trivial observation will map to all of $X$.
-Next, we must check arbitrary union and finite intersections.
-Recall that the experimental domain $\mathcal{S}$ has a countable basis $\mathcal{B} = \{\mathsf{o}_1,\mathsf{o}_2,\ldots\}$.
-
-Let $\{U_i\ |\ i\in I\}\subseteq \mathsf{T}$, where $I$ is some indexing set.
-Let $U_i = U(\mathsf{O}_i)$, where each $\mathsf{O}_i = \bigcup\limits_{j=1}^{\infty}\mathsf{o}_{n_{i,j}}$ for appropriately indexed basis elements $\mathsf{o}_{n_{i,j}}$, where $n_{i,j}$ is a positive integer.
-Now, since there are only countably many basis elements, we know that there are only countably many distinct $n_{i,j}$'s, so we may define a sequence of positive integers $m_k, k=1,2,\ldots$ such that $$\{m_k\}_{k=1}^{\infty} = \{n_{i,j}\ |\ i\in I, j=1,2,\ldots\}.$$ Then using Lemma \ref{ubehavior}, we have the following:
-$$
-\bigcup\limits_{i\in I}U(\mathsf{O}_i) = \bigcup\limits_{i\in I}\bigcup\limits_{j=1}^{\infty}U(\mathsf{o}_{n_{i,j}}) = \bigcup\limits_{k=1}^{\infty}U(\mathsf{o}_{m_k}) = U(\bigvee\limits_{k=1}^{\infty} \mathsf{o}_{m_k})
-$$
-which is an element of $\mathsf{T}$, so this collection is closed under arbitrary union.
-
-The proof for finite intersection is simpler. Let $\{U_i\}_{i=1}^N$ be a finite collection of elements of $\mathsf{T}$. Then as above writing $U_i = U(\mathsf{O}_i)$, we have:
-$$
-\bigcap\limits_{i=1}^N U(\mathsf{O}_i) = U(\bigwedge\limits_{i=1}^{N} \mathsf{O}_i)
-$$
-by Lemma \ref{ubehavior}, as required. Hence $\mathsf{T}$ is a topology. In what follows we will refer to elements of $\mathsf{T}$ as open sets.
-
-It is immediate that this topology is second-countable, since by Lemma \ref{ubehavior} the countable basis of observations for $\mathcal{S}$ becomes under $U$ a countable basis for the topology $\mathsf{T}$.
-To see why it is Hausdorff, let $x,y\in X$ be distinct possibilities.
-From Definition \ref{possibilitiesdef}, we have several cases to consider; however, it turns out that only one is possible.
-Without loss of generality, suppose $x\imp y$.
-But then we have $x\comp y$, so by symmetry we must also have $y\imp x$; in other words, $x=y$, contrary to assumption.
-It follows that $x\ncomp y$.
-Now suppose for a contradiction that for every open set $U\ni x$ and $V\ni y$ we have $U\cap V\neq \emptyset$.
-TODO: PROBLEM
-
+This is immediate from Lemma \ref{complements}.
 \end{proof}
 
+And so we have found that all experimentally distinguishable spaces may be given by Hausdorff, second-countable topological spaces.
+We end this section with an elementary result which has interesting consequences on the theory of experimentally distinguishable spaces.
+
 \begin{prop}
-	A Hausdorff and second countable space $X$ has at most cardinality of continuum.
+	A Hausdorff and second countable space $X$ has at most the cardinality of the continuum.
 \end{prop}
@@
 
-This means that, no matter what technique we use now or in the future, the collections of elements that we can properly define experimentally are at most infinite like the continuum.
+This means that, no matter what technique we use now or in the future, the collections of elements that we can properly define experimentally are at most the cardinality of the continuum.
 This already gives us a first simple and basic requirement a set of mathematical objects need to pass to be of scientific interest.
@@
 
-\section{Example of an Experimental Domain}
-
-Consider the collection of statements, ``this object has a mass strictly between $x$ and $y$ kilograms" where $x$ and $y$ are any positive real numbers with $x<y$.
-Consider the experimental test where we put the object on a balance with an appropriate level of precision for each corresponding statement.
-Then each observation may be equated to an open interval of positive real numbers $(x,y)\subset\mathbb{R}$.
-Conjunction and disjunction of these observations will correspond exactly to intersection and union of these intervals, respectively.
-Further, the collection of intervals of the form $(p,q)$ for $p,q\in\mathbb{Q}$ forms a basis of observations.
-Next, one can show that any sequence of observations satisfying the conditions of exact outcome corresponds precisely to a real number.
-In particular, the set of possibilities $X$ for this experimental domain will be the set of positive real numbers with their usual topology.
-The reader with some experience with elementary real analysis will recognize that this construction is essentially the same as the Cauchy sequence construction of the real numbers, where representative Cauchy sequences come from endpoints of intervals, and the intervals used those corresponding to observations in a sequence of observations yielding an exact outcome.
-
 \section{Experimental relationships}
@@
 Now that we have defined the possibilities of our domains and the topology associated with them, we want to establish relationships between them.
-We want to do this because the aim of scientific investigation is studying and characterizing these relationships
+We want to do this because the aim of scientific investigation is studying and characterizing these relationships.
 
@@
 
-TODO: add corollary environment for this one.
-
-\begin{prop}
+\begin{cor}
 The topology $\mathsf{T}(\mathcal{C})$ on the set of continuous functions $C(\mathbb{R},\mathbb{R})$ from $\mathbb{R}$ to itself is in general not equal to the open-open topology on $C(\mathbb{R},\mathbb{R})$.
-\end{prop}
+\end{cor}
 \begin{proof}
@@
 Funding for this work was provided in part by the MCubed program of the University of Michigan.
+The third author is supported by the National Science Foundation Graduate Research Fellowship Program under Grant No.
+DGE\#1256260.
+Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
 

### 2017-07-19 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Note added
Filling in the single degree of freedom

@@
 \section{Single degree of freedom}
-Define distribution over phase space.
 
-Show change under coordinate transformation.
+We first want to show how the math works in the simplest case, the one with a single degree of freedom.
+We will specifically use the simplest level of math for a number of reasons: it makes the details more accessible to a wider audience but it also shows that to show that the problem is not mathematically difficult once it is set up correctly.
 
-Define information entropy.
+Let's consider a normalized distribution over position and momentum $\rho(x, p)$.
+This can either be thought as a probability distribution or an actual physical distribution in phase space of an ideally continuous material.
+Suppose we have a differentiable law of evolution
+\begin{equation}
+\label{newCoordinates}
+\begin{aligned}
+\hat{q} = q + \frac{dq}{dt} dt  \\
+\hat{p} = p + \frac{dp}{dt} dt  \\
+\end{aligned}
+\end{equation}
+that moves each initial state $(q,p)$ to the final state $(\hat{q},\hat{p})$.
+Each little region $dq \,dp$ will contain $\rho(q, p) dq \,dp$ amount of material and will be mapped to a final region $d\hat{q} \,d\hat{p}$ that will contain $\rho(\hat{q}, \hat{p}) d\hat{q} \,d\hat{p}$ amount of material, which will have to match the initial amount.
 
-Show change under coordinate transformation.
+The Jacobian of the transformation is given by
+\begin{equation}
+\label{Jacobian}
+\begin{aligned}
+	|J| &= \left| \begin{matrix}
+		\dfrac{\partial \hat{q}}{\partial q} & \dfrac{\partial \hat{q}}{\partial p} \\[2.2ex]
+		\dfrac{\partial \hat{p}}{\partial q} & \dfrac{\partial \hat{p}}{\partial p} \end{matrix} \right| = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial x} \frac{\partial \hat{q}}{\partial p}\\
+	&= \{\hat{q}, \hat{p}\} \\
+	&= \left| \begin{matrix}
+	1 + \dfrac{\partial}{\partial x} \left( \dfrac{dx}{dt} \right) dt & \dfrac{\partial}{\partial p} \left( \dfrac{dx}{dt} \right) dt \\[2.2ex]
+	\dfrac{\partial}{\partial x} \left( \dfrac{dp}{dt} \right) dt & 1 + \dfrac{\partial}{\partial p} \left( \dfrac{dp}{dt} \right) dt \end{matrix} \right| \\
+&= 1 + \left[ \dfrac{\partial}{\partial x} \left( \dfrac{dx}{dt} \right) + \dfrac{\partial}{\partial p} \left( \dfrac{dp}{dt} \right) \right] dt + O(dt^2)\\
+\end{aligned}
+\end{equation}
+which coincides with the Poisson brackets $\{\hat{q}, \hat{p}\}$ of the new values for position and momentum.
 
-Show that conservation of information entropy is equivalent to transformation having Jacobian equals to 1.
+The transformation rules for the density and the area of phase space, then, are the following:
+\begin{equation}
+\label{newDistribution}
+\begin{aligned}
+d\hat{q}\,d\hat{p} &= |J| dq \,dp  \\
+\hat{\rho}(\hat{q}, \hat{p}) &= \frac{\rho(q, p)}{|J|}  \\
+\end{aligned}
+\end{equation}
 
-Show that canonical transformations are the ones and only the ones that have the Jacobian equals to one.
-Show relation to Poisson bracket.
-Show relation to incompressible flow.
-Show relation to symplectic form.
+Given a distribution, we can calculate its information entropy. On the initial distribution it will be given by:
+\begin{equation}
+\label{entropy}
+H(\rho) = -\int \rho \log \rho \; dx \,dp\\
+\end{equation}
+
+We can calculate how the entropy changes just by substituting \ref{newDistribution} in the definition for the entropy of the final distribution.
+\begin{equation}
+\label{newEntropy}
+\begin{aligned}
+H(\hat{\rho}) &= -\int \hat{\rho} \log \hat{\rho} \; d\hat{x}\,d\hat{p}= -\int \frac{\rho}{|J|} \log \frac{\rho}{|J|} \; |J| dx \,dp \\
+&= -\int \rho \log \frac{\rho}{|J|} \; dx \,dp \\
+&= -\int \rho \log \rho \; dx \,dp + \int \rho \log |J| \; dx \,dp \\
+\end{aligned}
+\end{equation}
+
+In other words, the final entropy is the initial entropy plus the expectation of the logarithm of the Jacobian.
+
+If the evolution obeys Hamilton's equations
+\begin{equation}
+\label{Hamilton}
+\begin{aligned}
+\frac{dq}{dt} = \frac{\partial H}{\partial p}  \\
+\frac{dp}{dt} = - \frac{\partial H}{\partial q}  \\
+\end{aligned}
+\end{equation}
+then we calculate the Jacobian
+\begin{equation}
+\label{HamiltonianJacobian}
+\begin{aligned}
+|J| &= 1 + \left[ \dfrac{\partial}{\partial q} \frac{\partial H}{\partial p} - \dfrac{\partial}{\partial p} \frac{\partial H}{\partial q} \right] dt + O(dt^2)\\
+&= 1 + O(dt^2)\\
+\end{aligned}
+\end{equation}
+which turns out to be unitary. Therefore the areas, the values of the distribution at each point and the information entropy are all conserved during the evolution.
+
+This should not be surprising.
+Hamiltonian mechanics leads to Liouville's theorem, which means preservation of phase space areas.
+Hamiltonian mechanics of canonical coordinates, therefore the new coordinates are also canonical and their Poisson bracket is unitary.
+Hamiltonian mechanics means invariance of the form $\omega$ with components
+\begin{equation}
+\label{singleForm}
+\omega_{ab} = TODO
+\end{equation}
+
+which corresponds to the two dimensional vector product (i.e.
+the area given by two vectors).
+Hamiltonian mechanics means incompressible flow in phase space, which means the value of the distribution has the same value for the initial and final state.
+And if the distribution is value by value the same, its entropy is going to be the same as well.
+All these are different aspect of the same property.
+
+We can easily run the argument backward. Suppose we start with a distribution over space and a law of evolution that preserves information entropy. How can we characterize the law?
+
+For entropy to be conserved, the logarithm of Jacobian must be zero everywhere, therefore the Jacobian must be unitary. This means that the law has to satisfy the condition
+\begin{equation}
+\label{divergenceFree}
+\dfrac{\partial}{\partial q} \left( \dfrac{dq}{dt} \right) + \dfrac{\partial}{\partial p} \left( \dfrac{dp}{dt} \right) = 0
+\end{equation}
+In other words: the vector field formed by the displacement vector $\left[ \dfrac{dq}{dt}, \dfrac{dp}{dt}\right]$ is divergence free, which means it admits a potential $H$ such that:
+\begin{equation*}
+\begin{aligned}
+\frac{dq}{dt} &= \frac{\partial H}{\partial p}  \\
+\frac{dp}{dt} &= - \frac{\partial H}{\partial q}
+\end{aligned}
+\end{equation*}
+So we have recovered Hamilton's equations.
+
+Therefore is not just that Hamiltonian mechanics on a single degree of freedom preserves information entropy: it is the only evolution that does. In fact, it's even deeper than that.
+
+As we saw, the information entropy is, in general, coordinate dependent.
+If we had a distribution over position only, we would not be able to arbitrarily change reference frame: information would not be invariant.
+A two dimensional manifold, with position and momentum, is exactly what is needed to define distributions such that their value and their information entropy is defined independent of reference frame.
+The symplectic form $\omega$ keeps track the areas on which these distributions are defined in a way similar to how the metric sensor $g$ keeps track about length in Riemannian geometry.
+
+Both phase space and Hamilton's equation are exactly the mathematical structures necessary and sufficient to describe distributions for which the information entropy is coordinate independent and is conserved in time.
 

### 2017-07-20 Articles/TopologyAndExperimentalDistinguishability/article.tex
made revisions for more mathematical writing style for several sections

@@
 
+\newcommand{\R}{\mathbb{R}}
+\newcommand{\Q}{\mathbb{Q}}
+\newcommand{\Z}{\mathbb{Z}}
+
 \theoremstyle{plain}% default 
@@
 \journal{Topology and its Applications}
-\bibliographystyle{elsarticle-num}
 
@@
 We then consider observations that identify elements in a set and show how they induce a Hausdorff and second-countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
-We then show that only continuous functions preserve experimental distinguishability and that the collection of these functions can be given a Hausdorff and second-countable topology.
-
+We then show that continuous functions preserve experimental distinguishability and define a Hausdorff and second-countable topology for this collection.
 \end{abstract}
@@
 \begin{keyword}
-topological spaces \sep TODO
-\MSC[2010] TODO \sep TODO
+topological spaces \sep scientific foundations \sep function space
+\MSC[2010] 54H99 \sep 00A79
 \end{keyword}
@@
 
-The successful use of mathematical ideas in experimental sciences is something long established and celebrated.\footnote{Unreasonable effectiveness and other} In particular topology is arguably the most widespread, as it provides the foundation for other mathematical tools such as differential geometry and ???.
-Why is it so successful?
+The successful use of mathematical ideas in experimental sciences is long established and celebrated \cite{wigner}.
+In many ways, topology is perhaps the most widespread, as it provides the foundation for other mathematical tools such as differential geometry and analysis.
+Topology is foundational in physics \cite{nakahara}, and has been the backbone of advances in many areas such as biology \cite{rashevsky} and social choice theory \cite{chichilnisky}.
+This leads one to ask, why is it so successful?
 What property is captured by topological spaces that is so fundamental for scientific investigation?
 
-We believe we have identified the answer: topologies capture the way objects are experimentally distinguishable.
-It is open sets that are associated with experimental outcomes, not the points themselves.
-For example, as real values are only measurable with finite precision, the standard topology on Euclidean spaces is the one that is physically meaningful.
-Along the same line, experimental relationships need to preserve what can be experimentally distinguished, and they are therefore represented as continuous functions.
+We attempt to answer these questions in a precise manner by defining the appropriate connections.
+Topological spaces generalize metric spaces; they provide a notion of ``nearness" of points without necessarily quantifying it.
+We extend this interpretation in a precise way to cover spaces representing objects (points in space, objects needing classification, etc.) which are distinguishable by scientific experimentation.
+Because experimental data is universally limited by precision, it is neighborhoods of points that are associated with experimental outcomes, as opposed to the points themselves.
+Along the same line, experimental relationships need to preserve what can be experimentally distinguished, and they are therefore should be realized as continuous functions on the appropriate topological spaces.
 
-In this sense, experimental distinguishability is \emph{the} application of topology in experimental science.
-It is fitting, then, that topology is so widespread given that defining what can be experimentally identified is truly a fundamental aspect of science.
+In this sense, experimental distinguishability is intrinsically topological in nature.
+It is fitting, then, that the use of topology is so widespread, as defining what can be experimentally identified is a fundamental aspect of science.
 The aim of this paper is to lay down a framework that formalizes this insight.
@@
 We first define experimental observations as statements paired with an experimental test that is able to verify them.
+These will be the fundamental objects which enable us to define our universe of discourse.
 We study their properties under logical operations and show that they are only closed under finite conjunction and countable disjunction.
-We then show that any set of objects that are experimentally distinguishable is a Hausdorff second-countable topological space.
-Both the points and the topology of the space are formally defined in terms of the observations themselves.
+We then show that any set of objects that are experimentally distinguishable is a Hausdorff, second-countable topological space.
+The topology of the space is formally defined in terms of the observations themselves.
 Finally, we study relationships between experimentally distinguishable objects.
@@
 
-We believe this work not only provides interesting insights, but a firm justification of the use of topological ideas in the sciences.
-
 \section{Experimental observations}
 
-In science, something is accepted as true only if one can provide a way to independently test the assertion. We introduce the following definitions to capture this notion.
+In science, a statement is considered valid for inquiry only if one can provide a way to independently test the assertion. We introduce the following definitions to capture this notion.
 
@@
 
-Given a statement, we wish to consider when it makes scentific sense to test. The following construction will define the building blocks of this theory.
+Given a statement, we wish to consider when it makes scentific sense to test. The following construction will define the next building block of this theory.
 
 \begin{defn}
-	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e. it can be restarted and stopped an arbitrary number of times) that, if successful, terminates in a finite amount of time.
+	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e.
+it can be restarted and stopped an arbitrary number of times) that either terminates in finite time (in which case it is said to be a successful experimental test), or continues indefinitely.
 \end{defn}
@@
 \begin{defn}
-	An \textbf{experimental observation} is a tuple $\mathsf{o} = \llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is always successful.
+	An \textbf{experimental observation} is a tuple $\llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$, normally denoted by $\mathsf{o}$, consisting of a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is successful.
 The experimental observation is \textbf{verified} if the statement is true.
@@
 \begin{defn}
-	An \textbf{experimental counter-observation} is a tuple $\mathsf{o}^c= \llbracket \mathsf{s}, \mathsf{e} \rrbracket$ composed by a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is false if and only if the experimental test is always successful.
+	An \textbf{experimental counter-observation} is a tuple $\llbracket \mathsf{s}, \mathsf{e} \rrbracket$, normally denoted by $\mathsf{o}^C$, consisting of a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is false if and only if the experimental test is always successful.
 The experimental counter-observation is \textbf{refuted} if the statement is false.
@@
 \begin{defn}
-	Given two experimental observations $\mathsf{o}_1$ and $\mathsf{o}_2$ we define:
+	Given two experimental observations $\mathsf{o}_1 = \llparenthesis \mathsf{s}_1,\mathsf{e}_1\rrparenthesis$ and $\mathsf{o}_2 = \llparenthesis \mathsf{s}_2,\mathsf{e}_2\rrparenthesis$ we define:
 	\begin{itemize}
@@
 
-Being able to test experimental observations one at a time is not enough.
-Given a set of observations we must be able to test them all to find those that are verified.
+In order to access arbitrary levels of precision in some system, we must be able to conduct arbitrarily many tests.
+Given a collection of observations $\mathcal{X}$, we wish to understand both the largest collection $\mathcal{Y}\supset\mathcal{X}$ of observations which can be deduced from $\mathcal{X}$, and similarly, we wish to find a small enough collection $\mathcal{Z}\subset\mathcal{X}$ from which we can deduce all of $\mathcal{X}$.
 With this in mind, we introduce the following definitions.
@@
 \begin{defn}
-	An \textbf{experimental domain} is a set of observations closed under finite conjunction and countable disjunction such that all observations can be tested in infinite time. 
+	An \textbf{experimental domain} is a set of observations closed under finite conjunction and countable disjunction, such that all observations can be tested in infinite time. 
 \end{defn}
@@
 Note that any infinite sub-basis will generate a basis of the same cardinality.
-We claim that it is only possible to verify all members of an experimental domain with a countable (sub-)basis even given infinite time.
+We claim that it is only possible to verify all members of an experimental domain with a (sub-)basis of at most countably infinite cardinality, even given infinite time.
 
@@
 If there exists a countable basis $\mathcal{B}$, then given infinite time one can test all observations in $\mathcal{B}$.
-Given all truth values for elements of the basis, one can deduce the truth values for the other observations in $\mathcal{S}$ (again using infinite time) by computing the appropriate conjunctions and disjunctions.
+Given which observations of the basis are verified, one can deduce which other observations in $\mathcal{S}$ are verified (again using infinite time) by computing the appropriate disjunctions.
 
@@
 \begin{rem}
-It is not hard to see that the (countable) disjunction of an entire basis produces an experimental observation which is always verified.
-Henceforth, we call this the \textbf{trivial observation}, denoted $\mathsf{o}_T$.
+If the (countable) disjunction of an entire basis produces an experimental observation which is always verified, we call this the \textbf{trivial observation}, denoted $\mathsf{o}_T$.
+This may not always be the case: consider, for example, the rather uninteresting case of an experimental domain consisting only of the null observation.
 \end{rem}
@@
 
-Given an experimental domain, we want to study the maximal sets of consistent observations that can be verified at the same time.
-These will define the possibilities of our domain: the possible cases that can be experimentally distinguished through the observations.
+Now that we have a framework for testing a system, we will define the class of systems we can understand with these tests.
+The framework will be general enough that one can readily apply it to most real-world experimental scenarios.
+In this construction, we will begin with a space we wish to understand - in this case, finding the appropriate element - and then define the appropriate collection of experimental observations to do so.
 
@@
 \end{defn}
+
+Given the possibility space $(X,x^*)$, we define the following distinguished subsets:
 
@@
 There exists an experimental observation $\mathsf{o} = (x^*\in U, \mathsf{e}_\in(U))$.
-Consider $\neg \mathsf{o} = (\neg x^*\in U, \mathsf{e}_\in(U)) = (x^*\notin U, \mathsf{e}_{\notin}(U^C)) = (x^*\in U^C, \mathsf{e}_{\notin}(U^C))$.
-That is, there exists a counter-observation $(x^*\in U^C, \mathsf{e}_{\notin}(U^C))$.
-Therefore $U^C$ is a refutable set.
+Consider $\neg \mathsf{o} = \neg(x^*\in U, \mathsf{e}_\in(U)) = (x^*\notin U, \mathsf{e}_{\notin}(U^C)) = (x^*\in U^C, \mathsf{e}_{\notin}(U^C))$.
+That is, there exists a counter-observation $(x^*\in U^C, \mathsf{e}_{\notin}(U^C))$, and so $U^C$ is a refutable set.
 \end{proof}
@@
 \begin{cor}
-The collection of refutable sets for $X$ is precisely the closed sets for the topology $\mathsf{T}$.
+The collection of refutable sets for $X$ is precisely the collection of closed sets for the topology $\mathsf{T}$.
 \end{cor}
@@
 
-TODO: the above propositions might generalize to any uncountable, second-countable space. This will be considered during the paper-writing process later.
-
 Thus we have shown one can start with two Hausdorff, second-countable spaces, and generate a new Hausdorff second-countable space consisting of all functions between them.
@@
 		\hline 
-		Hausdorff, second-countable topological space & Experimentally distinguishable space, whose points are the possible values and whose open sets represent the experimentally attainable levels of precision \\
+		$T_2$, 2nd ctbl space. & Experimentally distinguishable space, whose points are the possible values and whose open sets represent the experimentally attainable levels of precision \\
 		Open set & Verifiable set. We can verify experimentally that an element is within the set  \\ 
 		Closed set & Refutable set. We can verify experimentally that an element is not in the set \\ 
-		Basis of a topology & A collection of verifiable sets such that any verifiable set is determined by the basis sets\\
+		Basis of top. & A collection of verifiable sets such that any verifiable set is determined by the basis sets\\
 		Continuous \newline function &  A function between two sets of experimental distinguishable elements that preserves distinguishability \\
@@
 
+\bibliographystyle{elsarticle-num}
 \bibliography{bibliography}

### 2017-07-21 Articles/TopologyAndExperimentalDistinguishability/article.tex
updated first few sections in light of thursday meeting

@@
 
+Statements in this work shall have a truth value as well as \textit{logical content}, and so one can compare the meaning of statements as well as their truth values.
+For example, a statement can be a \textbf{contradiction}, which is stronger than being false.
+For example, the statement ``$\{1,2,3\}$ has four elements, and $\{1,2,3\}$ has three elements" is a contraction, which is a distinct claim from noticing it is a false statement.
+
 Given a statement, we wish to consider when it makes scentific sense to test. The following construction will define the next building block of this theory.
@@
 	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e.
-it can be restarted and stopped an arbitrary number of times) that either terminates in finite time (in which case it is said to be a successful experimental test), or continues indefinitely.
+it can be restarted and stopped an arbitrary number of times) which may be successful, in which case it terminates in finite time, or may be unsuccessful, in which case it may or may not terminate.
 \end{defn}
@@
 
-\begin{defn}
-	An \textbf{experimental counter-observation} is a tuple $\llbracket \mathsf{s}, \mathsf{e} \rrbracket$, normally denoted by $\mathsf{o}^C$, consisting of a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is false if and only if the experimental test is always successful.
-The experimental counter-observation is \textbf{refuted} if the statement is false.
-\end{defn}
-
-Experimental observations are the core object of the scientific discourse. We now define some basic relationships between experimental observations.
-
-\begin{defn}
-	Given two experimental observations $\mathsf{o}_1 = \llparenthesis \mathsf{s}_1,\mathsf{e}_1\rrparenthesis$ and $\mathsf{o}_2 = \llparenthesis \mathsf{s}_2,\mathsf{e}_2\rrparenthesis$ we define:
-	\begin{itemize}
-	\item \textbf{implication}: $\mathsf{o}_1 \imp \mathsf{o}_2$ ($\mathsf{o}_1$ implies $\mathsf{o}_2$) if $\mathsf{o}_2$ is always verified whenever $\mathsf{o}_1$ is verified
-	\item \textbf{equivalence}: $\mathsf{o}_1 = \mathsf{o}_2$ ($\mathsf{o}_1$ is equal to $\mathsf{o}_2$) if $\mathsf{o}_1 \imp \mathsf{o}_2$ and $\mathsf{o}_2 \imp \mathsf{o}_1$
-	\item \textbf{compatibility}: $\mathsf{o}_1 \comp \mathsf{o}_2$ ($\mathsf{o}_1$ is compatible with $\mathsf{o}_2$) if there exists a case in which $\mathsf{o}_1$ and $\mathsf{o}_2$ are jointly verified
-	\end{itemize}
-    The negation of these properties will be noted by $\nimp$, $\neq$ , $\ncomp$ respectively.
-\end{defn}
-
-A note on the language: even if $\mathsf{o}_1$ implies $\mathsf{o}_2$ this does not give us a way to verify $\mathsf{o}_2$.
-There may be cases where the $\mathsf{o}_2$ is verified and $\mathsf{o}_1$ is not.
-Therefore it is improper to say that one observation verifies another: only experimental tests verify observations.
-
-Lastly, we define a special type of observation.
-
-\begin{defn}
-The \textbf{null observation}, denoted $\mathsf{o}_N$, is any experimental observation that is never verified (i.e. the experimental test will never terminate successfully).
-\end{defn}
-
-The null observation is the degenerate case and has the following, maybe counterintuitive, properties.
-Since it is never verified, it can never rule out any other observations: the null observation implies all other observations.
-Since it is never verified, it can never be jointly verified with another observation: the null observation is incompatible with every experimental observation, including itself.
+Experimental observations are the core object of the scientific discourse.
+This construction is designed to encode the forming of a claim together with a procedure to test the claim.
+In this work, we do not presuppose knowledge of the truth values of the statements - only by ``running" the tests can we ascertain these values.
 
@@
 
-We now want to understand how experimental observations behave under logical operations. The resulting algebra will have some key differences from the standard Boolean algebra.
+We now want to understand how experimental observations behave under logical operations.
+While utilizing similar techniques, the resulting algebra will have some key differences from the standard Boolean algebra.
 
@@
 Taken as whole, they define the \emph{algebra of experimental observations}.
-With each definition, we will prove that the resulting object is an experimental (counter-)observation.
-
-\begin{defn}
-	The \textbf{negation} or \textbf{logical NOT} of an experimental observation $\mathsf{o}=\llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ is the experimental counter-observation $\neg \mathsf{o}=\llbracket\neg \mathsf{s}, \mathsf{e}\rrbracket$ where the negation of the statement $\mathsf{s}$ can be refuted by the experimental test $\mathsf{e}$.
-\end{defn}
-
-\begin{proof}
-	We show that the negation exists and is well defined.
-Let $\mathsf{o}=\llparenthesis \mathsf{s}, \mathsf{e}\rrparenthesis$ be an experimental observation.
-The experimental test $\mathsf{e}$ succeeds if and only if $\mathsf{s}$ is true or, equivalently, if and only if $\neg \mathsf{s}$ is false.
-So the tuple $\llbracket\neg \mathsf{s}, \mathsf{e}\rrbracket$ forms an experimental counter-observation.
-\end{proof}
+With each definition, we will prove that the resulting object is an experimental observation.
 
@@
 \begin{proof}
-	We show that the conjunction exists and is well defined. Let $\mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ be the experimental procedure defined as follows:
+	We show that the conjunction exists and is well defined. Let $\mathsf{e}_\wedge=\mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ be the experimental procedure defined as follows:
 	\begin{enumerate}
-	\item for each $i=1..n$ run the test $\mathsf{e}_i$
-	\item if all tests terminated successfully terminate successfully.
+	\item for each $i=1,\ldots,n$ run the test $\mathsf{e}_i$
+	\item if all tests $\mathsf{e}_i$ terminate successfully, terminate the test $\mathsf{e}_\wedge$ successfully.
 	\end{enumerate}
 	The experimental procedure so defined terminates successfully if and only if all $\mathsf{e}_i$ terminate successfully.
-It will do so in finite time as each $\mathsf{e}_i$ succeeds in finite time and there only a finite number of them.
-Therefore $\mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ is an experimental test that is successful if and only if all statements $\{\mathsf{s}_i\}_{i=1}^{n}$ are successful.
-So $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis\bigwedge\limits_{i=1}^{n} \mathsf{s}_i, \mathsf{e}_{\wedge}(\mathsf{e}_i)\rrparenthesis$ is an experimental observation.
+It will do so in finite time as each of the finitely many $\mathsf{e}_i$ succeeds in finite time.
+Therefore $\mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ is an experimental test that is successful if and only if all statements $\{\mathsf{s}_i\}_{i=1}^{n}$ are true.
+So, $\bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis\bigwedge\limits_{i=1}^{n} \mathsf{s}_i, \mathsf{e}_{\wedge}(\mathsf{e}_i)\rrparenthesis$ is an experimental observation.
 \end{proof}
@@
 \begin{proof}
-	We show that the disjunction exists and is well defined. Let $\mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ be the experimental procedure defined as follows:
+	We show that the disjunction exists and is well defined. Let $\mathsf{e}_\vee=\mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ be the experimental procedure defined as follows:
 	\begin{enumerate}
 	\item initialize $n$ to 1
-	\item for each $i=1..n$
+	\item for each $i=1,\ldots,n$:
 	\begin{enumerate}
 		\item run the test $\mathsf{e}_i$ for $n$ seconds
-		\item if $\mathsf{e}_i$ terminated successfully, terminate successfully
+		\item if $\mathsf{e}_i$ terminated successfully, terminate $\mathsf{e}_\vee$ successfully
 	\end{enumerate}
@@
 	\end{enumerate}
-	Suppose there exists an $i \in \mathbb{Z}^+$ such that $\mathsf{e}_i$ will terminate successfully, then the above procedure will eventually run that test for a time long enough and terminate successfully.
+	Suppose there exists an $i \in \mathbb{Z}^+$ such that $\mathsf{e}_i$ will terminate successfully.
+Then the above procedure will eventually run that test for time sufficient for it to terminate successfully.
 It will do so in finite time as it will have run finitely many tests finitely many times each for a finite amount of time.
@@
 
-\begin{rem}
-A useful equivalent characterization of $\mathsf{o}\imp\mathsf{o}'$ for observations $\mathsf{o}$ and $\mathsf{o}'$ is provided by the above constructions.
-We note that $\mathsf{o}\imp\mathsf{o}'$ if and only if $\mathsf{o}\wedge\mathsf{o}' = \mathsf{o}'$ if and only if $\mathsf{o}\vee\mathsf{o}'=\mathsf{o}$.
-\end{rem}
+Now, we define a special type of observation.
+
+\begin{defn}
+A \textbf{null observation}, denoted $\mathsf{o}_N$, is any experimental observation whose statement is a contradiction.
+\end{defn}
+
+We will need to compare certain observations with each other, and an important relationship is the following:
+\begin{defn}
+Two experimental observations $\mathsf{o}_1$ and $\mathsf{o}_2$ are said to be \textbf{incompatible} or \textbf{contradictory} if the conjunction $\mathsf{o}_1\wedge\mathsf{o}_2$ is a null observation.
+\end{defn}
 
@@
 
-In order to access arbitrary levels of precision in some system, we must be able to conduct arbitrarily many tests.
-Given a collection of observations $\mathcal{X}$, we wish to understand both the largest collection $\mathcal{Y}\supset\mathcal{X}$ of observations which can be deduced from $\mathcal{X}$, and similarly, we wish to find a small enough collection $\mathcal{Z}\subset\mathcal{X}$ from which we can deduce all of $\mathcal{X}$.
+Given a collection of observations $\mathcal{X}$, we wish to understand both the largest collection $\mathcal{Y}\supset\mathcal{X}$ of observations which can be constructed from $\mathcal{X}$ by our algebra of observations, and similarly, we wish to find smaller collections $\mathcal{Z}\subset\mathcal{X}$ from which we can construct all of $\mathcal{X}$.
 With this in mind, we introduce the following definitions.
@@
 \begin{prop}
-Let $\mathcal{S}$ be an experimental domain. Then there exists a countable basis (equivalently, sub-basis) $\mathcal{B}$ of $\mathcal{S}$.
+Let $\mathcal{D}$ be an experimental domain. Then there exists a countable basis (equivalently, sub-basis) $\mathcal{B}$ of $\mathcal{S}$.
 \end{prop}
@@
 If there exists a countable basis $\mathcal{B}$, then given infinite time one can test all observations in $\mathcal{B}$.
-Given which observations of the basis are verified, one can deduce which other observations in $\mathcal{S}$ are verified (again using infinite time) by computing the appropriate disjunctions.
+Given which observations of the basis are verified, one can deduce which other observations in $\mathcal{D}$ are verified (again using infinite time) by computing the appropriate disjunctions.
 
-If there does not exist a countable basis, then by definition there does not exist a sequence of experimental observations in $\mathcal{S}$ from which one can deduce all other observations in $\mathcal{S}$.
+If there does not exist a countable basis, then by definition there does not exist a sequence of experimental observations in $\mathcal{D}$ from which one can deduce all other observations in $\mathcal{D}$.
 Hence it is impossible to test all members of $\mathcal{S}$.
@@
 
-Now that we have a framework for testing a system, we will define the class of systems we can understand with these tests.
-The framework will be general enough that one can readily apply it to most real-world experimental scenarios.
+Now that we have a framework for testing a system, we will define a class of systems we can understand with these tests.
+The framework will be general enough that one can readily apply it to many real-world experimental scenarios.
 In this construction, we will begin with a space we wish to understand - in this case, finding the appropriate element - and then define the appropriate collection of experimental observations to do so.
+In what follows, let $X$ be a set.
 
 \begin{defn}
-A \textbf{set of possibilities} $(X,x^*)$ is a set together with a distinguished element $x^*$. 
-\end{defn}
-
-Given the possibility space $(X,x^*)$, we define the following distinguished subsets:
-
-\begin{defn}
-	A \textbf{verifiable set} $U \subseteq X$ is a subset of possibilities for which there exists an experimental observation $\mathsf{o} = (x^*\in U, \mathsf{e}_\in(U))$ where $\mathsf{e}_\in(U)$ is an experimental test that succeeds only if the object to identify is an element of $U$.
+For $x^*\in X$ and $U\subseteq X$, an \textbf{experimental identification} is any experimental observation of the form $\mathsf{o} = (x^*\in U, \mathsf{e}_\in(U))$ where $\mathsf{e}_\in(U)$ is an experimental test that succeeds if and only if $x^*\in U$.
+Any subset $U\subseteq X$ for which such an observation exists is said to be a \textbf{verifiable subset}.
 \end{defn}
@@
 \begin{defn}
-	A \textbf{refutable set} $U \subseteq X$ is a subset of possibilities for which there exists an experimental counter-observation $\mathsf{o} = (x^*\in U, \mathsf{e}_{\notin}(U))$ where $\mathsf{e}_{\notin}(U)$ is an experimental test that succeeds only if the object to identify is not an element of $U$.
+A \textbf{Experimental Identification Problem} $(X,\mathcal{D})$ is a set $X$ of cardinality $|X|\geq2$ together with a collection $\mathcal{D}$ of experimental identifications for the same element $x^*\in X$.
 \end{defn}
 
-\begin{lem}
-\label{complements}
-	The complement $U^C$ of a verifiable set $U \subseteq X$ is a refutable set.
-\end{lem}
-
-\begin{proof}
-	Let $U\subset X$ be a verifiable set.
-There exists an experimental observation $\mathsf{o} = (x^*\in U, \mathsf{e}_\in(U))$.
-Consider $\neg \mathsf{o} = \neg(x^*\in U, \mathsf{e}_\in(U)) = (x^*\notin U, \mathsf{e}_{\notin}(U^C)) = (x^*\in U^C, \mathsf{e}_{\notin}(U^C))$.
-That is, there exists a counter-observation $(x^*\in U^C, \mathsf{e}_{\notin}(U^C))$, and so $U^C$ is a refutable set.
-\end{proof}
+Next, we will explore the important properties of the collection of verifiable subsets.
 
@@
 
-Next, we wish to isolate the possibility spaces for which there exists an appropriate collection of experimental observations.
-Let $(X,x^*)$ be a set of experimental possibilities.
-An \textbf{experimental identification} is any experimental observation of the form $(x^*\in S,\mathsf{e}_{\in}(S))$, where $S$ is any subset of $X$.
+Next, we wish to isolate the experimental identification problems for which there exists an appropriate (for applications) collection of experimental observations.
 
 \begin{defn}
-	A set of experimental possibilities $X$ is \textbf{experimentally distinguishable} if there exists a collection $\mathsf{T}$ of experimental identifications which form an experimental domain such that given any two possibilities $x_1, x_2 \in X$ we can always find two mutually exclusive experimental observation such that each possibility is compatible with one observation.
+An experimental identification problem $(X,\mathcal{D})$ is called \textbf{experimentally distinguishable} if $\mathcal{D}$ is an experimental domain such that given any two possibilities $x_1, x_2 \in X$ we can find two experimental observations $(x^*\in U_1, \mathsf{e}_\in(U_1)), (x^*\in U_2, \mathsf{e}_\in(U_2))\in\mathcal{D}$ such that $x_i\in U_i$ for $i=1,2$, and $U_1\cap U_2 = \emptyset$.
 \end{defn}
@@
 \begin{thm}
-An experimentally distinguishable set $X$ together with its collection $\mathsf{T}$ of experimental identifications (realized as subsets of $X$) forms a Hausdorff, second-countable topology with open sets given by the verifiable sets.
+An experimentally distinguishable identification problem $(X,\mathcal{D})$ forms a Hausdorff, second-countable topology $(X,\mathsf{T})$ with open sets given by the verifiable sets. 
 \end{thm}
 \begin{proof}
-It is immediate that the trivial and null observations may be written as experimental identifications, and that they give $X\in\mathsf{T}$ and $\emptyset\in\mathsf{T}$.
-Proposition \ref{setbehavior} shows that the collection $\mathsf{T}$ is closed under finite intersection and countable union.
-Because $\mathsf{T}$ is given by an experimental domain of observations, there is a countable basis, so it is second-countable.
+First, from the definition one can see that for all $x\in X$, there exists a verifiable set $U$ with $x\in U$, so their union $X$ is also a verifiable set.
+Further, there is a null observation by taking the conjunction of two identifications corresponding to disjoint sets, so the empty set is a verifiable set.
+Now, Proposition \ref{setbehavior} shows that the collection $\mathsf{T}$ is closed under finite intersection and countable union.
+Because $\mathsf{T}$ is determined by an experimental domain of observations, there is a countable basis of operations which translates to a countable basis of open (verifiable) sets, so it is second-countable.
 To show it is closed under arbitrary union, notice that an arbitrary union may be rewritten as a union of basis elements, which is then a countable union, and so it remains in $\mathsf{T}$.
-\end{proof}
-
-\begin{cor}
-The collection of refutable sets for $X$ is precisely the collection of closed sets for the topology $\mathsf{T}$.
-\end{cor}
-\begin{proof}
-This is immediate from Lemma \ref{complements}.
+That the topology is Hausdorff is immediate from the last part of the definition.
 \end{proof}
@@
 And so we have found that all experimentally distinguishable spaces may be given by Hausdorff, second-countable topological spaces.
-We end this section with an elementary result which has interesting consequences on the theory of experimentally distinguishable spaces.
 
-\begin{prop}
-	A Hausdorff and second countable space $X$ has at most the cardinality of the continuum.
-\end{prop}
-
-\begin{proof}
-	We define an injective function $F:X\to2^{\mathbb{N}}$, where $2^{\mathbb{N}}$ denotes all infinite binary sequences.
-
-	Since $X$ is second-countable, we can enumerate the countable basis $\mathcal{B}$ as $\mathcal{B} = \{B_i\}_{i=1}^{\infty}$.
-Let $2^{\mathbb{N}}$ denote all infinite binary sequences.
-We define $F:X\to2^{\mathbb{N}}$ such that $F(x) = (F(x)_i)_{i=1}^{\infty}$ is the sequence where each element is given by:
-	$$
-	F(x)_i = 
-	\begin{cases}
-	1 & x\in B_i \\
-	0 & x\notin B_i
-	\end{cases}
-	$$
-	This is an injective function.
-Suppose $x_1 \neq x_2$, then since $X$ is Hausdorff there is at least one element of the basis that contains one but not the other.\footnote{In fact, $T_0$ separability would be sufficient.} Therefore $F(x_1) \neq F(x_2)$.
-As $F$ injects $X$ into $2^{\mathbb{N}}$, we have $|X| \leq |2^{\mathbb{N}}|=|\mathbb{R}|$.
-$X$ has at most cardinality of continuum.
-\end{proof}
-
+\begin{rem}
+It is an elementary exercise to show that any Hausdorff, second-countable topological space has cardinality at most that of the continuum.
 This means that, no matter what technique we use now or in the future, the collections of elements that we can properly define experimentally are at most the cardinality of the continuum.
@@
 \end{itemize}
+\end{rem}
 
@@
 
+TODO
+
 Now that we have defined the possibilities of our domains and the topology associated with them, we want to establish relationships between them.
-We want to do this because the aim of scientific investigation is studying and characterizing these relationships.
+The aim of scientific investigation is studying and characterizing these relationships.
 

### 2017-07-21 Articles/TopologyAndExperimentalDistinguishability/article.tex
possibly finished with core content revisions

@@
 In this construction, we will begin with a space we wish to understand - in this case, finding the appropriate element - and then define the appropriate collection of experimental observations to do so.
+The idea is that we have some object, and $X$ is a collection of possible attributes.
+We wish to identify the appropriate attribute among the collection.
+For example, the object might be an electron, and the collection of possible attributes might be the set of real numbers, representing possible masses.
 In what follows, let $X$ be a set.
@@
 \begin{thm}
-An experimentally distinguishable identification problem $(X,\mathcal{D})$ forms a Hausdorff, second-countable topology $(X,\mathsf{T})$ with open sets given by the verifiable sets. 
+An experimentally distinguishable identification problem $(X,\mathcal{D})$ has a natural Hausdorff, second-countable topology $(X,\mathsf{T})$ with open sets given by the verifiable sets. 
 \end{thm}
@@
 
+We refer to the topological space $(X,\mathsf{T})$ resulting from an experimentally distinguishable identification problem as an \textbf{experimentally distinguishable space}.
 And so we have found that all experimentally distinguishable spaces may be given by Hausdorff, second-countable topological spaces.
@@
 
-TODO
+Now that we have defined the possibilities of our domains and the topology associated with them, we want to establish relationships between experimentally distinguish.
+The aim of scientific investigation is studying and characterizing these relationships, and so we provide a formalism which not only gives a clear definition of experimental relationships, but we find that experimental relationships themselves can be given the structure of an experimentally distinguishable space.
 
-Now that we have defined the possibilities of our domains and the topology associated with them, we want to establish relationships between them.
-The aim of scientific investigation is studying and characterizing these relationships.
+First, suppose we are given a function $f:X\to Y$, where $Y$ is an experimentally distinguishable space $(Y,\mathsf{T}_Y)$ whose distinguished element is $y^*$, and suppose there exists $x^*\in X$ with $f(x)=y^*$.
+
+\begin{prop}
+For each $\mathsf{o}_Y = \llparenthesis y^*\in U_Y,\mathsf{e}_{\in}(U_Y)\rrparenthesis$, there exists an induced experimental identification $\mathsf{o}_X = \llparenthesis x^*\in f^{-1}(U_Y), \mathsf{e}_{\in}(f^{-1}(U_Y)) \rrparenthesis$.
+\end{prop}
+\begin{proof}
+We wish to show that $\mathsf{e}_{\in}(f^{-1}(U_Y))$ is an experimental test which succeeds if and only if $x^*\in f^{-1}(U_Y)$.
+To run this test, instead we run $\mathsf{e}_{\in}(U_Y)$ on $y^*$.
+We show first that if this test succeeds, then the statement $x^*\in f^{-1}(U_Y)$ is true.
+This test succeeds if and only if $y^*\in U_Y$.
+But $y^* \in U_Y$ if and only if $f^{-1}(\{y^*\})\subseteq f^{-1}(U_Y)$.
+Since $f(x^*)=y^*$, this implies $x^*\in f^{-1}(U_Y)$.
+
+Next, if $x^*\in f^{-1}(U_Y)$, then we have $y^*=f(x^*) \in U_Y$, and so the test $\mathsf{e}_{\in}(U_Y)$ succeeds.
+Hence, we have the observation $\llparenthesis x^*\in f^{-1}(U_Y), \mathsf{e}_{\in}(f^{-1}(U_Y))\rrparenthesis$.
+
+\end{proof}
+
+We call the induced observation the \textbf{pullback} $\mathsf{o}_X = f^*\mathsf{o}_Y$ of the observation $\mathsf{o}_Y$ via $f$.
+The following corollary is an immediate consequence, and gives the most useful interpretation of continuity in the context of experimentally distinguishable spaces.
+
+\begin{cor}
+Let $(X,\mathsf{T}_X)$ and $(Y,\mathsf{T}_Y)$ be experimentally distinguishable spaces with respective experimental domains $\mathcal{D}_X$ and $mathcal{D}_Y$, and let $f:X\to Y$ be a map with $f(x^*)=y^*$.
+Then $f$ is continuous if and only if all pullbacks of observations from $\mathcal{D}_Y$ are observations in $\mathcal{D}_X$.
+\end{cor}
+
+Next, we consider the opposite problem.
+Suppose we have two experimentally distinguishable spaces $(X,\mathsf{T}_X)$ and $(Y,\mathsf{T}_Y)$, and that we have a well-behaved (with respect to conjunction and disjunction) map on the experimental domains.
+Then this translates to a map of topologies, taking the verifiable (open) sets of $Y$ to those of $X$, which respects (arbitrary) union and (finite) intersection.
+In this case, we will show that one can find a continuous map which induces this map of topologies.
 
@@
 	\label{setfunctions}
-	Let $X$ and $Y$ be two Hausdorff topological spaces, and let $g: \mathsf{T}(Y) \rightarrow \mathsf{T}(X)$ be a mapping such that:
+	Let $X$ and $Y$ be two Hausdorff topological spaces, and let $g: \mathsf{T}_Y \rightarrow \mathsf{T}_X$ be a mapping such that:
 	\begin{enumerate}
-		\item It is compatible with union and intersection $\forall V_1, V_2 \in Y$ $g(V_1 \cup V_2)=g(V_1)\cup g(V_2)$ and $g(V_1 \cap V_2)=g(V_1)\cap g(V_2)$
+		\item It is compatible with union and intersection, i.e. for any subsets $V_1, V_2 \subseteq Y$, we have $g(V_1 \cup V_2)=g(V_1)\cup g(V_2)$ and $g(V_1 \cap V_2)=g(V_1)\cap g(V_2)$
 		\item $g(\emptyset) = \emptyset$
@@
 	\end{enumerate}
-	Then there exists a unique continuous function $f: X \rightarrow Y$ such that $g = f^{-1} |_{\mathsf{T}(Y)}$.
+	Then there exists a unique continuous function $f: X \rightarrow Y$ such that $g = f^{-1} |_{\mathsf{T}_Y}$.
 \end{prop}
@@
 \begin{proof}
-	We claim there exists a unique extension $\bar{g}:\sigma(Y)\to\sigma(X)$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma(X)$ and $\sigma(Y)$, such that $\bar{g}|_{\mathsf{T}(Y)}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
-Let $\bar{g}(V) = g(V)$ for all open sets $V \in \mathsf{T}(Y)$.
+	We claim there exists a unique extension $\bar{g}:\sigma(Y)\to\sigma(X)$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma(X)$ and $\sigma(Y)$, such that $\bar{g}|_{\mathsf{T}_Y}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
+Let $\bar{g}(V) = g(V)$ for all open sets $V \in \mathsf{T}_Y$.
 Let $A \in \sigma(Y)$ (not necessarily open) and $A^C$ be its complement.
@@
 
-For continuous functions to be physically distinguishable, we need to show they can always be given, as a set, given a topology that is Hausdorff and second-countable.
-
+For the collection of continuous functions to be experimentally distinguishable, we need a Hausdorff, second-countable topology.
 To that end, we introduce the basis-to-basis topology on the set of continuous functions from two topological spaces.

### 2017-07-27 Articles/TopologyAndExperimentalDistinguishability/article.tex
Revising topapp article before Christine

--- 2017-07-21 Articles/TopologyAndExperimentalDistinguishability/article.tex
@@
 
-In science, a statement is considered valid for inquiry only if one can provide a way to independently test the assertion. We introduce the following definitions to capture this notion.
+In science, a statement can be accepted as true only if there exists a way to independently verify it. To capture this notion we introduce the following definitions.
 
@@
 
-Statements in this work shall have a truth value as well as \textit{logical content}, and so one can compare the meaning of statements as well as their truth values.
-For example, a statement can be a \textbf{contradiction}, which is stronger than being false.
-For example, the statement ``$\{1,2,3\}$ has four elements, and $\{1,2,3\}$ has three elements" is a contraction, which is a distinct claim from noticing it is a false statement.
+\begin{defn}
+	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e.
+it can be restarted and stopped an arbitrary number of times) which may be successful, in which case it terminates in finite time, or may be not successful, in which case it may or may not terminate.
+\end{defn}
 
-Given a statement, we wish to consider when it makes scentific sense to test. The following construction will define the next building block of this theory.
+% https://ebooks.adelaide.edu.au/m/mill/john_stuart/system_of_logic/chapter18.html for black swan reference
 
 \begin{defn}
-	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e.
-it can be restarted and stopped an arbitrary number of times) which may be successful, in which case it terminates in finite time, or may be unsuccessful, in which case it may or may not terminate.
-\end{defn}
-
-\begin{defn}
-	An \textbf{experimental observation} is a tuple $\llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$, normally denoted by $\mathsf{o}$, consisting of a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is successful.
+	An \textbf{experimental observation} $\mathsf{o}$ is a tuple $\llparenthesis \mathsf{s}, \mathsf{e} \rrparenthesis$ consisting of a statement $\mathsf{s}$ and an experimental test $\mathsf{e}$ such that the statement is true if and only if the  experimental test is successful.
 The experimental observation is \textbf{verified} if the statement is true.
@@
 
-Experimental observations are the core object of the scientific discourse.
-This construction is designed to encode the forming of a claim together with a procedure to test the claim.
-In this work, we do not presuppose knowledge of the truth values of the statements - only by ``running" the tests can we ascertain these values.
-
 \section{Algebra of experimental observations}
 
-We now want to understand how experimental observations behave under logical operations.
-While utilizing similar techniques, the resulting algebra will have some key differences from the standard Boolean algebra.
+We now want to understand how the standard Boolean algebra defined on statements carries over to experimental observations.
 
-We define the following logical operations on experimental observations.
-Taken as whole, they define the \emph{algebra of experimental observations}.
-With each definition, we will prove that the resulting object is an experimental observation.
+\begin{rem}
+	Experimental observations are not closed under negation. The existence of an experimental test to verify a statement does not imply the existence of an experimental test to verify its negation.
+\end{rem}
 
@@
 \end{proof}
+
+\begin{rem}
+	Conjunction cannot be extended to a countable collection as verification would require infinite time.
+\end{rem}
 
@@
 
-Now, we define a special type of observation.
+Taken together, finite conjunction and countable disjunction form the algebra of experimental observations. We also introduce the following special case, which will be useful later.
 
 \begin{defn}
-A \textbf{null observation}, denoted $\mathsf{o}_N$, is any experimental observation whose statement is a contradiction.
+Any experimental observation whose statement is a contradiction is also called \textbf{contradiction} and is noted by $\bot$.
 \end{defn}
 
-We will need to compare certain observations with each other, and an important relationship is the following:
 \begin{defn}
-Two experimental observations $\mathsf{o}_1$ and $\mathsf{o}_2$ are said to be \textbf{incompatible} or \textbf{contradictory} if the conjunction $\mathsf{o}_1\wedge\mathsf{o}_2$ is a null observation.
+Two experimental observations $\mathsf{o}_1$ and $\mathsf{o}_2$ are said to be \textbf{incompatible} if the conjunction $\mathsf{o}_1\wedge\mathsf{o}_2$ is a contradiction.
 \end{defn}
@@
 
-Given a collection of observations $\mathcal{X}$, we wish to understand both the largest collection $\mathcal{Y}\supset\mathcal{X}$ of observations which can be constructed from $\mathcal{X}$ by our algebra of observations, and similarly, we wish to find smaller collections $\mathcal{Z}\subset\mathcal{X}$ from which we can construct all of $\mathcal{X}$.
-With this in mind, we introduce the following definitions.
+We now want to characterize the sets of observations for which it is feasible to experimentally verify all true statements.
 
@@
 \end{defn}
+
+\begin{rem}
+	We do allow infinite time for the verification of a domain with the understanding that some domains will only be approximately verified in finite time.
+As we have, so to speak, only one infinity to spend, we spend it here to maximize its usefulness.
+\end{rem}
+
+At this point, the similarities between this mathematical structure and topologies are starting to emerge. In analogy to the latter, we define the following.
 
@@
 
-Clearly, given a sub-basis one can generate a basis by taking all finite conjunctions of the observations.
-Note that any infinite sub-basis will generate a basis of the same cardinality.
-We claim that it is only possible to verify all members of an experimental domain with a (sub-)basis of at most countably infinite cardinality, even given infinite time.
+\begin{rem}
+	As for topologies, given a sub-basis one can generate a basis by taking all finite conjunctions of the observations. Any infinite sub-basis will generate a basis of the same cardinality.
+\end{rem}
 
@@
 
-\begin{rem}
-If the (countable) disjunction of an entire basis produces an experimental observation which is always verified, we call this the \textbf{trivial observation}, denoted $\mathsf{o}_T$.
-This may not always be the case: consider, for example, the rather uninteresting case of an experimental domain consisting only of the null observation.
-\end{rem}
-
 \section{Experimental distinguishability}
 
-Now that we have a framework for testing a system, we will define a class of systems we can understand with these tests.
-The framework will be general enough that one can readily apply it to many real-world experimental scenarios.
-In this construction, we will begin with a space we wish to understand - in this case, finding the appropriate element - and then define the appropriate collection of experimental observations to do so.
-In what follows, let $X$ be a set.
+We now turn our attention to a more specific case. We want to characterize an experimental domain whose function is to identify the value of an attribute among all possible values.
 
 \begin{defn}
-For $x^*\in X$ and $U\subseteq X$, an \textbf{experimental identification} is any experimental observation of the form $\mathsf{o} = (x^*\in U, \mathsf{e}_\in(U))$ where $\mathsf{e}_\in(U)$ is an experimental test that succeeds if and only if $x^*\in U$.
-Any subset $U\subseteq X$ for which such an observation exists is said to be a \textbf{verifiable subset}.
+	An \textbf{attribute} is the triple $(X, x, \mathcal{D}_x)$ where:
+	\begin{itemize}
+		\item $X$ is the set of all possible values, which we call \textbf{possibilities}, and satisfies $|X|>1$
+		\item $x$ is the value of the attribute, therefore $x \in X$
+		\item $\mathcal{D}_x$ is an experimental domain containing all possible experimental observations of the form $\mathsf{o} = (x\in U, \mathsf{e}_\in(U))$ where $U \subseteq X$ is a set of possibilities and $\mathsf{e}_\in(U)$ is an experimental test that succeeds if and only if $x \in U$
+	\end{itemize}
+	Any subset $U\subseteq X$ for which such an observation exists is said to be a \textbf{verifiable set}.
 \end{defn}
-
-\begin{defn}
-A \textbf{Experimental Identification Problem} $(X,\mathcal{D})$ is a set $X$ of cardinality $|X|\geq2$ together with a collection $\mathcal{D}$ of experimental identifications for the same element $x^*\in X$.
-\end{defn}
-
-Next, we will explore the important properties of the collection of verifiable subsets.
 
@@
 , U_n \subseteq X$ be n verifiable sets.
-For each $U_i$ there exists an experimental observation $\mathsf{o}_i = (x^*\in U_i, \mathsf{e}_\in(U_i))$.
-Consider $\mathsf{o} = \bigwedge\limits_{i=1}^{n} \mathsf{o}_i = (\bigwedge\limits_{i=1}^{n} x^*\in U_i , \mathsf{e}_{\wedge}(\mathsf{e}_\in(U_i)))=( x^*\in \bigcap\limits_{i=1}^{n} U_i, \mathsf{e}_\in(\bigcap\limits_{i=1}^{n} U_i))$ is an experimental observation.
+For each $U_i$ there exists an experimental observation $\mathsf{o}_i = (x\in U_i, \mathsf{e}_\in(U_i))$.
+Consider $\mathsf{o} = \bigwedge\limits_{i=1}^{n} \mathsf{o}_i = (\bigwedge\limits_{i=1}^{n} x\in U_i , \mathsf{e}_{\wedge}(\mathsf{e}_\in(U_i)))=( x\in \bigcap\limits_{i=1}^{n} U_i, \mathsf{e}_\in(\bigcap\limits_{i=1}^{n} U_i))$ is an experimental observation.
 We conclude $\bigcap\limits_{i=1}^{n} U_i$ is a verifiable set.
@@
 \subseteq X$ be an infinite sequence of verifiable sets.
-For each $U_i$ there exists an experimental observation $\mathsf{o}_i = (x^*\in U_i, \mathsf{e}_\in(U_i))$.
-Consider $\mathsf{o} = \bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = (\bigvee\limits_{i=1}^{\infty} x^*\in U_i, \mathsf{e}_{\vee}(\mathsf{e}_\in(U_i)))=( x^*\in \bigcup\limits_{i=1}^{\infty} U_i, \mathsf{e}_\in(\bigcup\limits_{i=1}^{\infty} U_i))$ is an experimental observation.
+For each $U_i$ there exists an experimental observation $\mathsf{o}_i = (x\in U_i, \mathsf{e}_\in(U_i))$.
+Consider $\mathsf{o} = \bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = (\bigvee\limits_{i=1}^{\infty} x\in U_i, \mathsf{e}_{\vee}(\mathsf{e}_\in(U_i)))=( x\in \bigcup\limits_{i=1}^{\infty} U_i, \mathsf{e}_\in(\bigcup\limits_{i=1}^{\infty} U_i))$ is an experimental observation.
 Thus $\bigcup\limits_{i=1}^{\infty} U_i$ is a verifiable set.
@@
 \begin{defn}
-An experimental identification problem $(X,\mathcal{D})$ is called \textbf{experimentally distinguishable} if $\mathcal{D}$ is an experimental domain such that given any two possibilities $x_1, x_2 \in X$ we can find two experimental observations $(x^*\in U_1, \mathsf{e}_\in(U_1)), (x^*\in U_2, \mathsf{e}_\in(U_2))\in\mathcal{D}$ such that $x_i\in U_i$ for $i=1,2$, and $U_1\cap U_2 = \emptyset$.
+The possibilities of an attribute $(X, x, \mathcal{D}_x)$, and by extension the attribute itself, are said \textbf{experimentally distinguishable} if $\mathcal{D}_x$ is such that given any two possibilities $x_1, x_2 \in X$ we can find two incompatible experimental observations $(x\in U_1, \mathsf{e}_\in(U_1)), (x\in U_2, \mathsf{e}_\in(U_2))\in\mathcal{D}$ such that $x_i\in U_i$ for $i=1,2$.
 \end{defn}
 
-From the work above, we will quickly arrive at the following.
+We are now ready to prove the first main result of this work.
 
 \begin{thm}
-An experimentally distinguishable identification problem $(X,\mathcal{D})$ forms a Hausdorff, second-countable topology $(X,\mathsf{T})$ with open sets given by the verifiable sets. 
+The set of experimentally distinguishable possibilities for an attribute $(X, x, \mathcal{D}_x)$ is a Hausdorff, second-countable topological space $(X,\mathsf{T})$ where the open sets are given by the verifiable sets.
 \end{thm}
@@
 
-And so we have found that all experimentally distinguishable spaces may be given by Hausdorff, second-countable topological spaces.
-
 \begin{rem}
-It is an elementary exercise to show that any Hausdorff, second-countable topological space has cardinality at most that of the continuum.
-This means that, no matter what technique we use now or in the future, the collections of elements that we can properly define experimentally are at most the cardinality of the continuum.
-This already gives us a first simple and basic requirement a set of mathematical objects need to pass to be of scientific interest.
-
-For example, these objects have cardinality of continuum, and therefore are good candidates:
-\begin{itemize}
-	\item Euclidean space $\mathbb{R}^n$
-	\item all continuous functions from $\mathbb{R}$ to $\mathbb{R}$
-	\item all open sets in $\mathbb{R}^n$
-	\item all subsets of $\mathbb{N}$
-\end{itemize}
-
-These, instead, have cardinality greater then continuum, and therefore are not good candidates:
-\begin{itemize}
-	\item all functions from $\mathbb{R}$ to $\mathbb{R}$
-	\item all subsets of $\mathbb{R}$
-\end{itemize}
+As any Hausdorff, second-countable topological space has at most cardinality of the continuum, that is also the greatest cardinality that a set of experimentally distinguishable objects can have.
+We can conclude that set of mathematical objects, such as all functions from $\mathbb{R}$ to $\mathbb{R}$, that do not satisfy this requirement are not good candidates to represent scientific concepts.
 \end{rem}
@@
 
-TODO
+We now want to characterize relationships between two attributes.
+Such relationships can be defined either on the values (i.e.
+the possibilities) or on the observations (i.e.
+the experimental domain).
+We need to show that both definitions lead to the same mathematical object.
 
-Now that we have defined the possibilities of our domains and the topology associated with them, we want to establish relationships between them.
-The aim of scientific investigation is studying and characterizing these relationships.
+\begin{defn}[Experimental relationship on possibilities]
+	Let $(X, x, \mathcal{D}_x)$ and $(Y, y, \mathcal{D}_y)$ be two attributes. An \textbf{experimental relationship} is a map $f : X \rightarrow Y$ that can be used within an experimental test.
+\end{defn}
 
 \begin{prop}
-	\label{setfunctions}
-	Let $X$ and $Y$ be two Hausdorff topological spaces, and let $g: \mathsf{T}(Y) \rightarrow \mathsf{T}(X)$ be a mapping such that:
+	The experimental relationship $f$ defined above is a continuous function.
+\end{prop}
+\begin{proof}
+Let $o_y = \llparenthesis y\in U_Y,\mathsf{e}_{\in}(U_Y)\rrparenthesis \in \mathcal{D}_y$.  Consider the following experimental procedure:
 \begin{enumerate}
-		\item It is compatible with union and intersection $\forall V_1, V_2 \in Y$ $g(V_1 \cup V_2)=g(V_1)\cup g(V_2)$ and $g(V_1 \cap V_2)=g(V_1)\cap g(V_2)$
+	\item map $x$ to $y=f(x)$
+	\item run the test $\mathsf{e}_{\in}(U_Y)$
+\end{enumerate}
+It will be successful if and only if $y \in U_Y$.
+Since $y=f(x)$, it is successful if and only if $x \in f^{-1}(U_Y)$: the procedure is the test $\mathsf{e}_{\in}(f^{-1}(U_Y))$.
+This means that $o_x = \llparenthesis x \in f^{-1}(U_Y),\mathsf{e}_{\in}(f^{-1}(U_Y))\rrparenthesis$ is an experimental observation and it must be in $\mathcal{D}_x$ since $\mathcal{D}_x$ contains all possible experimental observations of that form.
+It follows that $f^{-1}(U_Y)$ is a verifiable set and must be part of the topology $\mathsf{T}_X$.
+\end{proof}
+
+Next, we consider the opposite problem.
+Suppose we have two experimentally distinguishable spaces $(X,\mathsf{T}_X)$ and $(Y,\mathsf{T}_Y)$, and that we have a well-behaved (with respect to conjunction and disjunction) map on the experimental domains.
+Then this translates to a map of topologies, taking the verifiable (open) sets of $Y$ to those of $X$, which respects (arbitrary) union and (finite) intersection.
+In this case, we will show that one can find a continuous map which induces this map of topologies.
+
+\begin{defn}[Experimental relationship on observations]
+	Let $(X, x, \mathcal{D}_x)$ and $(Y, y, \mathcal{D}_y)$ be two attributes.
+An \textbf{experimental relationship} is a map $g : \mathcal{D}_y \rightarrow \mathcal{D}_x$ such that if $\mathsf{o}_y \in \mathcal{D}_y$ is verified then $\mathsf{o}_x \in \mathcal{D}_x$ must also be verified.
+To be consistent, such relationship must have these properties:
+	\begin{enumerate}
+	\item it is compatible with conjunction and disjunction: for any $\mathsf{o}_1, \mathsf{o}_2 \in \mathcal{D}_y$, we have $g(\mathsf{o}_1 \wedge \mathsf{o}_2)=g(\mathsf{o}_1)\wedge g(\mathsf{o}_2)$ and $g(\mathsf{o}_1 \vee \mathsf{o}_2)=g(\mathsf{o}_1)\vee g(\mathsf{o}_2)$
+	\item contradiction leads to contradiction: $g(\bot) = \bot$
+	\item no knowledge leads to no knowledge: if $Y$ is the verifiable set associated with $\mathsf{o}_y$ then $X$ is the verifiable set associated with  $g(\mathsf{o}_y)$
+	\end{enumerate}
+\end{defn}
+
+\begin{prop}
+	Each experimental relationship $g$ defined above is associated with a unique continuous function $f: X \rightarrow Y$ such that $g(\llparenthesis y\in U_Y,\mathsf{e}_{\in}(U_Y)\rrparenthesis) = \llparenthesis x\in f^{-1}(U_Y),\mathsf{e}_{\in}(f^{-1}(U_Y))\rrparenthesis$.
+\end{prop}
+
+\begin{proof}
+TODO: 	Let $X$ and $Y$ be two Hausdorff topological spaces, and let $g: \mathsf{T}_Y \rightarrow \mathsf{T}_X$ be a mapping such that:
+\begin{enumerate}
+	\item It is compatible with union and intersection, i.e. for any subsets $V_1, V_2 \subseteq Y$, we have $g(V_1 \cup V_2)=g(V_1)\cup g(V_2)$ and $g(V_1 \cap V_2)=g(V_1)\cap g(V_2)$
 	\item $g(\emptyset) = \emptyset$
@@
 \end{enumerate}
-	Then there exists a unique continuous function $f: X \rightarrow Y$ such that $g = f^{-1} |_{\mathsf{T}(Y)}$.
-\end{prop}
 
-\begin{proof}
-	We claim there exists a unique extension $\bar{g}:\sigma(Y)\to\sigma(X)$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma(X)$ and $\sigma(Y)$, such that $\bar{g}|_{\mathsf{T}(Y)}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
-Let $\bar{g}(V) = g(V)$ for all open sets $V \in \mathsf{T}(Y)$.
+	We claim there exists a unique extension $\bar{g}:\sigma(Y)\to\sigma(X)$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma(X)$ and $\sigma(Y)$, such that $\bar{g}|_{\mathsf{T}_Y}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
+Let $\bar{g}(V) = g(V)$ for all open sets $V \in \mathsf{T}_Y$.
 Let $A \in \sigma(Y)$ (not necessarily open) and $A^C$ be its complement.
@@
 
-The above work gives us an approach to reconstruct continuous functions given the behavior of the inverse on open sets.
-In the context of collecting information on observable phenomena, the function $g$ represents the total of all information possible to gather on the correlation between two variables.
-The fact that from $g$ we can construct a unique continuous $f$ shows us that in the infinite-resource ideal, we fully obtain the information to give exact relations between variables.
-Further, the result itself tells us that the interesting phenomena in our framework of experimental observations will be continuous, which is often a baseline assumption in physics.
+We can now state the second main result of this work.
 
-For continuous functions to be physically distinguishable, we need to show they can always be given, as a set, given a topology that is Hausdorff and second-countable.
+\begin{thm}
+	An experimental relationship between two sets $X$ and $Y$ of experimentally distinguishable possibilities is a continuous function $f : X \rightarrow Y$ with respect to the between the respective topological spaces $(X,\mathsf{T}_X)$ and $(Y,\mathsf{T}_Y)$ given by the verifiable sets.
+\end{thm}
 
-To that end, we introduce the basis-to-basis topology on the set of continuous functions from two topological spaces.
-This is the topology generated by the sets of functions that map a basis element of one element inside an element of the other space.
+\begin{rem}
+	This result gives a formal justification as to why continuous functions are prevalent in science in general and in physics in particular.
+As topologies capture experimental distinguishability, continuous function preserve it.
+\end{rem}
+
+\section{Distinguishability of experimental relationships}
+
+To conclude we want to make sure that experimental relationships are themselves experimentally distinguishable.
+To do so it suffices to show that the set of continuous functions between two Hausdorff, second-countable topological spaces can be given a topology that is Hausdorff and second-countable.
 
@@
 
-We will also show that this is not in general equal to the open-open topology.
+\begin{rem}
+	Note that the basis-to-basis topology is not in general equal to the open-open topology. The former may depend on the basis chosen while the second is uniquely defined by the topology.
+\end{rem}
 
-\begin{prop}
-	In general, the topology $\mathsf{T}(\mathcal{C})$ defined above depends on the bases for $X$ and $Y$. 
-\end{prop}
-\begin{proof}
-	We will give an example of such a case.
-Let $X=Y=\mathbb{R}$ with the usual topology on $\mathbb{R}$.
-Let $\mathcal{B}_1$ be the basis for $\mathbb{R}$ consisting of all open intervals with rational endpoints, and let $\mathcal{B}_2 = \mathcal{B}_1\cup\{(0,\pi)\}$.
-We find a set open in our topology $\mathsf{T}(\mathcal{C})$ using $\mathcal{B}_2$ as our basis which is not open when we use $\mathcal{B}_1$.
-Consider the following set, open in our topology when using $\mathcal{B}_2$:
-	$$
-	V = \{f\in\mathcal{C}| f((0,\pi))\subset(0,1)\}
-	$$
-	This is clearly open because $(0,\pi)$ and $(0,1)$ are basis elements in $\mathcal{B}_2$.
-
-	Next, we show that when using $\mathcal{B}_1$ as our basis for $X=Y=\mathbb{R}$, no finite intersection and/or infinite union of sub-basis elements for $\mathsf{T}(\mathcal{C})$ will equal $V$.
-Henceforth, when we say ``open" subsets, unless otherwise stated, we will refer exclusively to the topology $\mathsf{T}(\mathcal{C})$ when using $\mathcal{B}_1$ for $\mathbb{R}$.
-The smallest sub-basis subsets containing $V$ are those of the form:
-	$$
-	B_r = \{f\in\mathcal{C} | f((0,r))\subset(0,1)\}
-	$$
-	for each $r\in\mathbb{Q}$ with $r<\pi$.
-By ``smallest" we mean any sub-basis elements containing $V$ which are not in the form of $B_r$ will contain some $B_r$.
-Now, any finite intersection of the $B_r$'s will be another $B_r$, but these strictly contain $V$, so we can never obtain $V$ from this sub-basis.
-
-	Similarly, the largest sub-basis sets contained within $V$ are of the form: 
-	$$
-	S_r = \{f\in\mathcal{C} | f((0,r))\subset(0,1)\}
-	$$
-	for each $r\in \mathbb{Q}$ with $r>\pi$.
-Let $S = \cup_{r>\pi}S_r$.
-This is the largest open set in our topology which is contained in $V$.
-Consider the function $f(x) = x/\pi$.
-This is continuous and is an element of $V$.
-But notice $f(\pi)=1$, so $f\notin S_r$ for all $r>\pi$, hence $f\notin S$.
-Thus, $V$ is a set which is open in $\mathsf{T}(\mathcal{C})$ when we use $\mathcal{B}_1$ for $X$ and $Y$, but not open when we use $\mathcal{B}_2$ for $X$ and $Y$.
-\end{proof}
-
-\begin{cor}
-The topology $\mathsf{T}(\mathcal{C})$ on the set of continuous functions $C(\mathbb{R},\mathbb{R})$ from $\mathbb{R}$ to itself is in general not equal to the open-open topology on $C(\mathbb{R},\mathbb{R})$.
-\end{cor}
-\begin{proof}
-In the proof of the previous proposition, note that the basis for the open-open topology is independent of the basis used for the underlying spaces, and in particular is a superset of the basis for $\mathsf{T}(\mathcal{C})$ when using $\mathbb{B}_2$ to construct it.
-Hence the open-open topology is distinct from $\mathsf{T}(\mathcal{C})$ when using the basis $\mathcal{B}_1$ on $\mathbb{R}$.
-\end{proof}
-
-Thus we have shown one can start with two Hausdorff, second-countable spaces, and generate a new Hausdorff second-countable space consisting of all functions between them.
-Topological spaces with these properties make up our class of scientifically interesting spaces, so the space of continuous (distinguishability-preserving) functions between spaces of distinguishable quantities is again a physically distinguishable space.
-Because one can iterate this construction (since the relevant properties are preserved), this is a way to generate arbitrarily many new physically distinguishable spaces encoding important information about the ``lower-order" spaces (i.e.
-they consist of functions which encode a conversion between verifiable sets in different spaces).
+As experimental relationships are themselves distinguishable, we can recursively form experimental relationships between experimental relationship leading to functions of arbitrary order while remaining within the definition provided.
+The framework is therefore complete.
 
@@
 	\end{tabular} 
-	\caption{Topology to physics dictionary}
+	\caption{Topology to physics dictionary. This table sums up the relationship established between math and scientific concepts.}
 \end{table}

### 2017-07-27 Articles/TopologyAndExperimentalDistinguishability/article.tex
Incorporating Christine's comments in the main body

@@
 
-In science, a statement can be accepted as true only if there exists a way to independently verify it. To capture this notion we introduce the following definitions.
+In science, a statement can be accepted as true only if there exists a way to experimentally verify it. To capture this notion we introduce the following definitions.
 
@@
 \begin{defn}
-Any experimental observation whose statement is a contradiction is also called \textbf{contradiction} and is noted by $\bot$.
+Any experimental observation whose statement is a contradiction is also called a \textbf{contradiction} and is noted by $\bot$.
 \end{defn}
@@
 \begin{rem}
-	We do allow infinite time for the verification of a domain with the understanding that some domains will only be approximately verified in finite time.
-As we have, so to speak, only one infinity to spend, we spend it here to maximize its usefulness.
+	We do allow infinite time for the verification of a domain with the understanding that some domains will only be partially verified in finite time.
+As we have, so to speak, only one infinity to spend, we spend it here to maximize its utility.
 \end{rem}
@@
 	A \textbf{sub-basis} of an experimental domain is any subset that can generate all others via finite conjunction and countable disjunction.
-A \textbf{basis} of an experimental domain is any subset that can generate all others by countable disjunctions.
+A \textbf{basis} of an experimental domain is any subset that can generate all others by countable disjunction.
 \end{defn}
@@
 \begin{rem}
-	As for topologies, given a sub-basis one can generate a basis by taking all finite conjunctions of the observations. Any infinite sub-basis will generate a basis of the same cardinality.
+	As for topologies, given a sub-basis one can generate a basis by taking all finite conjunctions. Any infinite sub-basis will generate a basis of the same cardinality.
 \end{rem}
@@
 \begin{prop}
-Let $\mathcal{D}$ be an experimental domain. Then there exists a countable basis (equivalently, sub-basis) $\mathcal{B}$ of $\mathcal{S}$.
+Let $\mathcal{D}$ be an experimental domain. Then there exists a countable basis (equivalently, sub-basis) $\mathcal{B}$ of $\mathcal{D}$.
 \end{prop}
@@
 If there does not exist a countable basis, then by definition there does not exist a sequence of experimental observations in $\mathcal{D}$ from which one can deduce all other observations in $\mathcal{D}$.
-Hence it is impossible to test all members of $\mathcal{S}$.
+Hence it is impossible to test all members of $\mathcal{D}$.
 \end{proof}
@@
 
-We now turn our attention to a more specific case. We want to characterize an experimental domain whose function is to identify the value of an attribute among all possible values.
+We now turn our attention to a more specific case. We want to characterize an experimental domain whose purpose is to identify an element among a set of possibilities.
 
 \begin{defn}
-	An \textbf{attribute} is the triple $(X, x, \mathcal{D}_x)$ where:
+	An \textbf{experimental identification domain} is the triplet $(\mathcal{D}_x, X, x)$ where:
 	\begin{itemize}
-		\item $X$ is the set of all possible values, which we call \textbf{possibilities}, and satisfies $|X|>1$
-		\item $x$ is the value of the attribute, therefore $x \in X$
+		\item $X$ is the set of \textbf{possibilities}, and satisfies $|X|>1$
+		\item $x$ is the element to identify among the possibilities, therefore $x \in X$
 		\item $\mathcal{D}_x$ is an experimental domain containing all possible experimental observations of the form $\mathsf{o} = (x\in U, \mathsf{e}_\in(U))$ where $U \subseteq X$ is a set of possibilities and $\mathsf{e}_\in(U)$ is an experimental test that succeeds if and only if $x \in U$
@@
 	Let $U_1, U_2, ...
-, U_n, ...$ be a countable infinite sequence of verifiable sets.
+, U_n, ...$ be a countably infinite sequence of verifiable sets.
 The finite intersection $\bigcap\limits_{i=1}^{n} U_i$ and the countable union $\bigcup\limits_{i=1}^{\infty} U_i$ are verifiable sets.
@@
 \begin{defn}
-The possibilities of an attribute $(X, x, \mathcal{D}_x)$, and by extension the attribute itself, are said \textbf{experimentally distinguishable} if $\mathcal{D}_x$ is such that given any two possibilities $x_1, x_2 \in X$ we can find two incompatible experimental observations $(x\in U_1, \mathsf{e}_\in(U_1)), (x\in U_2, \mathsf{e}_\in(U_2))\in\mathcal{D}$ such that $x_i\in U_i$ for $i=1,2$.
+A set of possibilities $X$ is \textbf{experimentally distinguishable} if there exists an experimental identification domain $(\mathcal{D}_x, X, x)$ such that for any two possibilities $x_1, x_2 \in X$ we can find two incompatible experimental observations $(x\in U_1, \mathsf{e}_\in(U_1)), (x\in U_2, \mathsf{e}_\in(U_2))\in\mathcal{D}$ such that $x_i\in U_i$ for $i=1,2$.
 \end{defn}
@@
 \begin{thm}
-The set of experimentally distinguishable possibilities for an attribute $(X, x, \mathcal{D}_x)$ is a Hausdorff, second-countable topological space $(X,\mathsf{T})$ where the open sets are given by the verifiable sets.
+A set of experimentally distinguishable possibilities $X$ has a natural Hausdorff, second-countable topology $(X,\mathsf{T})$ with the open sets given by the verifiable sets of the associated experimental identification domain $(\mathcal{D}_x, X, x)$.
 \end{thm}
@@
 As any Hausdorff, second-countable topological space has at most cardinality of the continuum, that is also the greatest cardinality that a set of experimentally distinguishable objects can have.
-We can conclude that set of mathematical objects, such as all functions from $\mathbb{R}$ to $\mathbb{R}$, that do not satisfy this requirement are not good candidates to represent scientific concepts.
+We can conclude that sets of mathematical objects, such as all functions from $\mathbb{R}$ to $\mathbb{R}$, that do not satisfy this requirement are not good candidates to represent scientific concepts.
 \end{rem}
@@
 
-We now want to characterize relationships between two attributes.
-Such relationships can be defined either on the values (i.e.
-the possibilities) or on the observations (i.e.
-the experimental domain).
+We now want to characterize relationships between two experimentally distinguishable elements.
+Such relationships can be defined either on the values (i.e.~the possibilities) or on the observations (i.e.~the experimental domain).
 We need to show that both definitions lead to the same mathematical object.
 
-\begin{defn}[Experimental relationship on possibilities]
-	Let $(X, x, \mathcal{D}_x)$ and $(Y, y, \mathcal{D}_y)$ be two attributes. An \textbf{experimental relationship} is a map $f : X \rightarrow Y$ that can be used within an experimental test.
+\begin{defn}[Experimental relationship between possibilities]
+	Let $(\mathcal{D}_x, X, x)$ and $(\mathcal{D}_y, Y, y)$ be two experimental identification domains.
+An \textbf{experimental relationship} is a map $f : X \rightarrow Y$ that can be used within an experimental test.
 \end{defn}
@@
 
-Next, we consider the opposite problem.
-Suppose we have two experimentally distinguishable spaces $(X,\mathsf{T}_X)$ and $(Y,\mathsf{T}_Y)$, and that we have a well-behaved (with respect to conjunction and disjunction) map on the experimental domains.
-Then this translates to a map of topologies, taking the verifiable (open) sets of $Y$ to those of $X$, which respects (arbitrary) union and (finite) intersection.
-In this case, we will show that one can find a continuous map which induces this map of topologies.
-
-\begin{defn}[Experimental relationship on observations]
-	Let $(X, x, \mathcal{D}_x)$ and $(Y, y, \mathcal{D}_y)$ be two attributes.
+\begin{defn}[Experimental relationship between observations]
+	Let $(\mathcal{D}_x, X, x)$ and $(\mathcal{D}_y, Y, y)$ be two experimental identification domains.
 An \textbf{experimental relationship} is a map $g : \mathcal{D}_y \rightarrow \mathcal{D}_x$ such that if $\mathsf{o}_y \in \mathcal{D}_y$ is verified then $\mathsf{o}_x \in \mathcal{D}_x$ must also be verified.
@@
 \begin{prop}
-	Each experimental relationship $g$ defined above is associated with a unique continuous function $f: X \rightarrow Y$ such that $g(\llparenthesis y\in U_Y,\mathsf{e}_{\in}(U_Y)\rrparenthesis) = \llparenthesis x\in f^{-1}(U_Y),\mathsf{e}_{\in}(f^{-1}(U_Y))\rrparenthesis$.
+	For each experimental relationship $g$ defined above there exists a unique continuous function $f: X \rightarrow Y$ such that $g(\llparenthesis y\in U_Y,\mathsf{e}_{\in}(U_Y)\rrparenthesis) = \llparenthesis x\in f^{-1}(U_Y),\mathsf{e}_{\in}(f^{-1}(U_Y))\rrparenthesis$.
 \end{prop}
@@
 \begin{thm}
-	An experimental relationship between two sets $X$ and $Y$ of experimentally distinguishable possibilities is a continuous function $f : X \rightarrow Y$ with respect to the between the respective topological spaces $(X,\mathsf{T}_X)$ and $(Y,\mathsf{T}_Y)$ given by the verifiable sets.
+	An experimental relationship between two sets $X$ and $Y$ of experimentally distinguishable possibilities is a continuous function $f : X \rightarrow Y$ between the respective natural topologies $(X,\mathsf{T}_X)$ and $(Y,\mathsf{T}_Y)$.
 \end{thm}
@@
 	This result gives a formal justification as to why continuous functions are prevalent in science in general and in physics in particular.
-As topologies capture experimental distinguishability, continuous function preserve it.
+As topologies capture experimental distinguishability, continuous functions preserve it.
 \end{rem}
@@
 Let $\mathcal{B}_X$ and $\mathcal{B}_Y$ be two bases for $X$ and $Y$ respectively.
-The basis-to-basis topology $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ on $C(X,Y)$ with respect to the basis $\mathcal{B}_X$ and $\mathcal{B}_Y$ is the topology generated by all sets of the form
+The \textbf{basis-to-basis topology} $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ on $C(X,Y)$ with respect to the basis $\mathcal{B}_X$ and $\mathcal{B}_Y$ is the topology generated by all sets of the form
 	$$
@@
 \end{defn}
-
-Now we need to show that if the two spaces $X$ and $Y$ are Hausdorff and second-countable, the basis-to-basis topology is Hausdorff and second-countable for a suitable choice of basis.
 
@@
 \begin{rem}
-	Note that the basis-to-basis topology is not in general equal to the open-open topology. The former may depend on the basis chosen while the second is uniquely defined by the topology.
+	Note that the basis-to-basis topology is not in general equal to the open-open topology.
+The former may depend on the choice of bases $\mathcal{B}_X$ and $\mathcal{B}_Y$ while the second is uniquely defined by the topologies of $X$ and $Y$.
 \end{rem}
 
-As experimental relationships are themselves distinguishable, we can recursively form experimental relationships between experimental relationship leading to functions of arbitrary order while remaining within the definition provided.
+As experimental relationships are themselves distinguishable, we can recursively form experimental relationships between experimental relationships leading to functions of arbitrary order while remaining within the definitions provided.
 The framework is therefore complete.

### 2017-07-27 Articles/TopologyAndExperimentalDistinguishability/article.tex
Revised introduction/conclusion and black swan

@@
 The successful use of mathematical ideas in experimental sciences is long established and celebrated \cite{wigner}.
-In many ways, topology is perhaps the most widespread, as it provides the foundation for other mathematical tools such as differential geometry and analysis.
-Topology is foundational in physics \cite{nakahara}, and has been the backbone of advances in many areas such as biology \cite{rashevsky} and social choice theory \cite{chichilnisky}.
+Topology is perhaps the most widespread, either directly\cite{rashevsky, chichilnisky} or as a foundation to other tools\cite{nakahara}.
 This leads one to ask, why is it so successful?
@@
 
-We attempt to answer these questions in a precise manner by defining the appropriate connections.
-Topological spaces generalize metric spaces; they provide a notion of ``nearness" of points without necessarily quantifying it.
-We extend this interpretation in a precise way to cover spaces representing objects (points in space, objects needing classification, etc.) which are distinguishable by scientific experimentation.
-Because experimental data is universally limited by precision, it is neighborhoods of points that are associated with experimental outcomes, as opposed to the points themselves.
-Along the same line, experimental relationships need to preserve what can be experimentally distinguished, and they are therefore should be realized as continuous functions on the appropriate topological spaces.
+We believe experimentally distinguishability to be the correct concept.
+The notion of ``nearness" captured by topologies keeps track of how hard it is to tell two elements apart scientific observation.
+The standard topology on a real line, for examples, captures the inability to measure a continuous value with infinite precision.
+This connection holds in cases where a metric would not be physically meaningful (e.g.
+phase space) and in discrete spaces (where each element can be individually identified).
+It is fitting, then, that the use of topology is so widespread, as defining what can be experimentally distinguished is a fundamental aspect of science.
 
-In this sense, experimental distinguishability is intrinsically topological in nature.
-It is fitting, then, that the use of topology is so widespread, as defining what can be experimentally identified is a fundamental aspect of science.
 The aim of this paper is to lay down a framework that formalizes this insight.
-
-The work is organized as follows.
 We first define experimental observations as statements paired with an experimental test that is able to verify them.
-These will be the fundamental objects which enable us to define our universe of discourse.
-We study their properties under logical operations and show that they are only closed under finite conjunction and countable disjunction.
-We then show that any set of objects that are experimentally distinguishable is a Hausdorff, second-countable topological space.
-The topology of the space is formally defined in terms of the observations themselves.
-Finally, we study relationships between experimentally distinguishable objects.
-These are represented by continuous functions, and they can be shown to be themselves physically distinguishable.
+We then study their properties under logical operations and conclude that they are only closed under finite conjunction and countable disjunction.
+Next we define experimental distinguishability and show, as our fist main result, that any set of objects that are experimentally distinguishable is a Hausdorff, second-countable topological space, where the topology is formally defined in terms of the observations themselves.
+Last we define experimental relationships between experimentally distinguishable objects and show, as our second main result, that experimental relationships are represented by continuous functions and are themselves physically distinguishable.
 
@@
 	An \textbf{experimental test} $\mathsf{e}$ is a repeatable procedure (i.e.
-it can be restarted and stopped an arbitrary number of times) which may be successful, in which case it terminates in finite time, or may be not successful, in which case it may or may not terminate.
+it can be restarted and stopped an arbitrary number of times) which may be successful, in which case it terminates in finite time, or may be not successful, in which case it may or may not terminate.\footnote{In line with philosophical tradition\cite{mill,russell}, we can define this experimental test:
+		\begin{enumerate}
+			\item find a swan
+			\item if black terminate successfully
+			\item go to step 1
+	\end{enumerate}}
 \end{defn}
-
-% https://ebooks.adelaide.edu.au/m/mill/john_stuart/system_of_logic/chapter18.html for black swan reference
 
@@
 These definitions lead naturally to topological spaces, giving the result that any set of experimentally distinguishable possibilities is a Hausdorff and second-countable topological space.
-In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects, as they too can be given a Hausdorff and secound-countable topology.
+In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects, as they too can be given a Hausdorff and second-countable topology.
 The ability to create functions of any order confirms that the universe of discourse provided by these definitions is closed and self-consistent.

### 2017-07-28 Articles/TopologyAndExperimentalDistinguishability/article.tex
Reviewed abstract and conclusions

@@
 \begin{abstract}
-We formalize the relationship between topological spaces and the ability to distinguish objects experimentally, providing understanding and justification as to why topological spaces and continuous functions are pervasive tools in science.
+	In this work we introduce the idea that the primary application of topology in experimental sciences is to keep track of what can be distinguished through experimentation.
+This link provides understanding and justification as to why topological spaces and continuous functions are pervasive tools in science.
 We first define an experimental observation as a statement that can be verified using an experimental procedure and show that observations are closed under finite conjunction and countable disjunction.
-We then consider observations that identify elements in a set and show how they induce a Hausdorff and second-countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
-We then show that continuous functions preserve experimental distinguishability and define a Hausdorff and second-countable topology for this collection.
+We then consider observations that identify elements within a set and show how they induce a Hausdorff and second-countable topology on that set, thus identifying an open set as one that can be associated with an experimental observation.
+We then show that experimental relationships are continuous functions, as they must preserve experimental distinguishability, and that they are themselves experimentally distinguishable by defining a Hausdorff and second-countable topology for this collection.
 \end{abstract}
@@
 
-We believe experimentally distinguishability to be the correct concept.
-The notion of ``nearness" captured by topologies keeps track of how hard it is to tell two elements apart scientific observation.
-The standard topology on a real line, for examples, captures the inability to measure a continuous value with infinite precision.
-This connection holds in cases where a metric would not be physically meaningful (e.g.
-phase space) and in discrete spaces (where each element can be individually identified).
+We believe experimental distinguishability to be the relevant concept.
+The notion of ``nearness" captured by topologies keeps track of how hard it is to tell two elements apart via scientific observation.
+The standard topology on a real line, for example, captures the inability to measure a continuous value with infinite precision.
+This connection holds in cases where a metric would not be physically meaningful (e.g.~phase space) and in discrete spaces (where each element can be individually identified).
 It is fitting, then, that the use of topology is so widespread, as defining what can be experimentally distinguished is a fundamental aspect of science.
@@
 Next we define experimental distinguishability and show, as our fist main result, that any set of objects that are experimentally distinguishable is a Hausdorff, second-countable topological space, where the topology is formally defined in terms of the observations themselves.
-Last we define experimental relationships between experimentally distinguishable objects and show, as our second main result, that experimental relationships are represented by continuous functions and are themselves physically distinguishable.
+Last we define experimental relationships between experimentally distinguishable objects and show, as our second main result, that experimental relationships are represented by continuous functions and are themselves experimentally distinguishable.
 
@@
 		\begin{enumerate}
-			\item find a swan
-			\item if black terminate successfully
-			\item go to step 1
+			\item Find a swan
+			\item If black terminate successfully
+			\item Go to step 1
 	\end{enumerate}}
@@
 	\end{tabular} 
-	\caption{Topology to physics dictionary. This table sums up the relationship established between math and scientific concepts.}
+	\caption{Topology to physics dictionary. This table sums up the relationship established between mathematical and scientific concepts.}
 \end{table}
@@
 
-We have shown that it is possible to formally capture the requirements of experimental sciences using experimental observations and their properties.
-These definitions lead naturally to topological spaces, giving the result that any set of experimentally distinguishable possibilities is a Hausdorff and second-countable topological space.
-In the same framework, experimental relationships between these objects are represented mathematically by continuous functions and are themselves experimentally distinguishable objects, as they too can be given a Hausdorff and second-countable topology.
-The ability to create functions of any order confirms that the universe of discourse provided by these definitions is closed and self-consistent.
+What emerges from this work is that \emph{the primary application of topology in science is experimental distinguishability}.
+Its role is to track of what can be distinguished through experimentation.
+The importance of continuous functions in science and physics is explained since experimental relationships have to be consistent with what can be experimentally distinguished.
+This provides a strong justification as to why deterministic systems are always characterized by continuous transformations.
+It is not that they happen to be continuous: they must be.
 

### 2017-07-28 Articles/TopologyAndExperimentalDistinguishability/article.tex
Notation and style of proofs is more consistent

@@
 \begin{proof}
-	We show that the conjunction exists and is well defined. Let $\mathsf{e}_\wedge=\mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ be the experimental procedure defined as follows:
+	To show that the conjunction is well defined it suffices to show that we can construct a suitable experimental test.
+Let $\mathsf{e}_\wedge=\mathsf{e}_\wedge(\{\mathsf{e}_i\}_{i=1}^{n})$ be the experimental procedure defined as follows:
 	\begin{enumerate}
@@
 	\end{enumerate}
-	The experimental procedure so defined terminates successfully if and only if all $\mathsf{e}_i$ terminate successfully.
+	This experimental procedure terminates successfully if and only if all $\mathsf{e}_i$ terminate successfully.
 It will do so in finite time as each of the finitely many $\mathsf{e}_i$ succeeds in finite time.
@@
 \begin{proof}
-	We show that the disjunction exists and is well defined. Let $\mathsf{e}_\vee=\mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ be the experimental procedure defined as follows:
+	As before, it suffices to show that we can construct a suitable experimental test.
+Let $\mathsf{e}_\vee=\mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ be the experimental procedure defined as follows:
 	\begin{enumerate}
@@
 \begin{defn}
-	An \textbf{experimental identification domain} is the triplet $(\mathcal{D}_x, X, x)$ where:
+	An \textbf{experimental identification domain} is the triplet $(\mathcal{D}_X, X, x)$ where:
 	\begin{itemize}
@@
 		\item $x$ is the element to identify among the possibilities, therefore $x \in X$
-		\item $\mathcal{D}_x$ is an experimental domain containing all possible experimental observations of the form $\mathsf{o} = (x\in U, \mathsf{e}_\in(U))$ where $U \subseteq X$ is a set of possibilities and $\mathsf{e}_\in(U)$ is an experimental test that succeeds if and only if $x \in U$
+		\item $\mathcal{D}_X$ is an experimental domain containing all possible experimental observations of the form $\mathsf{o} = \llparenthesis x\in U, \mathsf{e}_\in(U)\rrparenthesis$ where $U \subseteq X$ is a set of possibilities and $\mathsf{e}_\in(U)$ is an experimental test that succeeds if and only if $x \in U$
 	\end{itemize}
@@
 \begin{proof}
-	We claim the finite intersection of verifiable sets is a verifiable set.
+	We show that the finite intersection of verifiable sets is a verifiable set.
 Let $U_1, U_2, ...
 , U_n \subseteq X$ be n verifiable sets.
-For each $U_i$ there exists an experimental observation $\mathsf{o}_i = (x\in U_i, \mathsf{e}_\in(U_i))$.
-Consider $\mathsf{o} = \bigwedge\limits_{i=1}^{n} \mathsf{o}_i = (\bigwedge\limits_{i=1}^{n} x\in U_i , \mathsf{e}_{\wedge}(\mathsf{e}_\in(U_i)))=( x\in \bigcap\limits_{i=1}^{n} U_i, \mathsf{e}_\in(\bigcap\limits_{i=1}^{n} U_i))$ is an experimental observation.
-We conclude $\bigcap\limits_{i=1}^{n} U_i$ is a verifiable set.
+For each $U_i$ there exists an experimental observation $\mathsf{o}_i = \llparenthesis x\in U_i, \mathsf{e}_\in(U_i) \rrparenthesis$.
+Consider the finite conjunction $\mathsf{o} = \bigwedge\limits_{i=1}^{n} \mathsf{o}_i = \llparenthesis \bigwedge\limits_{i=1}^{n} x\in U_i , \mathsf{e}_{\wedge}(\mathsf{e}_\in(U_i)) \rrparenthesis=\llparenthesis x\in \bigcap\limits_{i=1}^{n} U_i, \mathsf{e}_\in(\bigcap\limits_{i=1}^{n} U_i)\rrparenthesis$ which is an experimental observation.
+Thus $\bigcap\limits_{i=1}^{n} U_i$ is a verifiable set.
 
-	We claim the countable union of verifiable sets is a verifiable set.
+	We show that the countable union of verifiable sets is a verifiable set.
 Let $U_1, U_2, ...
@@
 \subseteq X$ be an infinite sequence of verifiable sets.
-For each $U_i$ there exists an experimental observation $\mathsf{o}_i = (x\in U_i, \mathsf{e}_\in(U_i))$.
-Consider $\mathsf{o} = \bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = (\bigvee\limits_{i=1}^{\infty} x\in U_i, \mathsf{e}_{\vee}(\mathsf{e}_\in(U_i)))=( x\in \bigcup\limits_{i=1}^{\infty} U_i, \mathsf{e}_\in(\bigcup\limits_{i=1}^{\infty} U_i))$ is an experimental observation.
+For each $U_i$ there exists an experimental observation $\mathsf{o}_i = \llparenthesis x\in U_i, \mathsf{e}_\in(U_i)\rrparenthesis$.
+Consider the infinite disjunction $\mathsf{o} = \bigvee\limits_{i=1}^{\infty} \mathsf{o}_i = \llparenthesis\bigvee\limits_{i=1}^{\infty} x\in U_i, \mathsf{e}_{\vee}(\mathsf{e}_\in(U_i))\rrparenthesis=\llparenthesis x\in \bigcup\limits_{i=1}^{\infty} U_i, \mathsf{e}_\in(\bigcup\limits_{i=1}^{\infty} U_i)\rrparenthesis$ which is an experimental observation.
 Thus $\bigcup\limits_{i=1}^{\infty} U_i$ is a verifiable set.
@@
 
-Next, we wish to isolate the experimental identification problems for which there exists an appropriate (for applications) collection of experimental observations.
+To make sure we have enough experimental observations to tell the possibilities apart, we introduce the following definition.
 
 \begin{defn}
-A set of possibilities $X$ is \textbf{experimentally distinguishable} if there exists an experimental identification domain $(\mathcal{D}_x, X, x)$ such that for any two possibilities $x_1, x_2 \in X$ we can find two incompatible experimental observations $(x\in U_1, \mathsf{e}_\in(U_1)), (x\in U_2, \mathsf{e}_\in(U_2))\in\mathcal{D}$ such that $x_i\in U_i$ for $i=1,2$.
+A set of possibilities $X$ is \textbf{experimentally distinguishable} if there exists an experimental identification domain $(\mathcal{D}_X, X, x)$ such that for any two possibilities $x_1, x_2 \in X$ we can find two incompatible experimental observations $\llparenthesis x\in U_1, \mathsf{e}_\in(U_1)\rrparenthesis, \llparenthesis x\in U_2, \mathsf{e}_\in(U_2)\rrparenthesis\in\mathcal{D}_X$ such that $x_i\in U_i$ for $i=1,2$.
 \end{defn}
@@
 \begin{thm}
-A set of experimentally distinguishable possibilities $X$ has a natural Hausdorff, second-countable topology $(X,\mathsf{T})$ with the open sets given by the verifiable sets of the associated experimental identification domain $(\mathcal{D}_x, X, x)$.
+A set of experimentally distinguishable possibilities $X$ has a natural Hausdorff, second-countable topology $(X,\mathsf{T})$ with the open sets given by the verifiable sets of the associated experimental identification domain $(\mathcal{D}_X, X, x)$.
 \end{thm}
 \begin{proof}
-First, from the definition one can see that for all $x\in X$, there exists a verifiable set $U$ with $x\in U$, so their union $X$ is also a verifiable set.
-Further, there is a null observation by taking the conjunction of two identifications corresponding to disjoint sets, so the empty set is a verifiable set.
-Now, Proposition \ref{setbehavior} shows that the collection $\mathsf{T}$ is closed under finite intersection and countable union.
+First, from the definition one can see that for all $x\in X$, there exists a verifiable set $U$ with $x\in U$.
+Therefore the union of a (countable) basis is the verifiable set $X$.
+Further, there exist at least two incompatible experimental observation, corresponding to two disjoint sets, so the empty set is a verifiable set.
+Now, \ref{setbehavior} shows that the collection $\mathsf{T}$ is closed under finite intersection and countable union.
 Because $\mathsf{T}$ is determined by an experimental domain of observations, there is a countable basis of operations which translates to a countable basis of open (verifiable) sets, so it is second-countable.
@@
 \begin{defn}[Experimental relationship between possibilities]
-	Let $(\mathcal{D}_x, X, x)$ and $(\mathcal{D}_y, Y, y)$ be two experimental identification domains.
+	Let $(\mathcal{D}_X, X, x)$ and $(\mathcal{D}_X, Y, y)$ be two experimental identification domains.
 An \textbf{experimental relationship} is a map $f : X \rightarrow Y$ that can be used within an experimental test.
@@
 \begin{proof}
-Let $o_y = \llparenthesis y\in U_Y,\mathsf{e}_{\in}(U_Y)\rrparenthesis \in \mathcal{D}_y$.  Consider the following experimental procedure:
+Let $o_y = \llparenthesis y\in U_Y,\mathsf{e}_{\in}(U_Y)\rrparenthesis \in \mathcal{D}_Y$.  Since $f$ can be used within an experimental test, consider the following experimental procedure:
 \begin{enumerate}
@@
 Since $y=f(x)$, it is successful if and only if $x \in f^{-1}(U_Y)$: the procedure is the test $\mathsf{e}_{\in}(f^{-1}(U_Y))$.
-This means that $o_x = \llparenthesis x \in f^{-1}(U_Y),\mathsf{e}_{\in}(f^{-1}(U_Y))\rrparenthesis$ is an experimental observation and it must be in $\mathcal{D}_x$ since $\mathcal{D}_x$ contains all possible experimental observations of that form.
+This means that $o_x = \llparenthesis x \in f^{-1}(U_Y),\mathsf{e}_{\in}(f^{-1}(U_Y))\rrparenthesis$ is an experimental observation and it must be in $\mathcal{D}_X$ since $\mathcal{D}_X$ contains all possible experimental observations of that form.
 It follows that $f^{-1}(U_Y)$ is a verifiable set and must be part of the topology $\mathsf{T}_X$.
@@
 \begin{defn}[Experimental relationship between observations]
-	Let $(\mathcal{D}_x, X, x)$ and $(\mathcal{D}_y, Y, y)$ be two experimental identification domains.
-An \textbf{experimental relationship} is a map $g : \mathcal{D}_y \rightarrow \mathcal{D}_x$ such that if $\mathsf{o}_y \in \mathcal{D}_y$ is verified then $\mathsf{o}_x \in \mathcal{D}_x$ must also be verified.
+	Let $(\mathcal{D}_X, X, x)$ and $(\mathcal{D}_Y, Y, y)$ be two experimental identification domains.
+An \textbf{experimental relationship} is a map $g : \mathcal{D}_Y \rightarrow \mathcal{D}_X$ such that if $\mathsf{o} \in \mathcal{D}_Y$ is verified then $g(\mathsf{o}) \in \mathcal{D}_X$ must also be verified.
 To be consistent, such relationship must have these properties:
@@
 	\item contradiction leads to contradiction: $g(\bot) = \bot$
-	\item no knowledge leads to no knowledge: if $Y$ is the verifiable set associated with $\mathsf{o}_y$ then $X$ is the verifiable set associated with  $g(\mathsf{o}_y)$
+	\item no knowledge leads to no knowledge: if $Y$ is the verifiable set associated with $\mathsf{o} \in \mathcal{D}_Y$ then $X$ is the verifiable set associated with  $g(\mathsf{o})$
 	\end{enumerate}
@@
 \begin{proof}
-TODO: 	Let $X$ and $Y$ be two Hausdorff topological spaces, and let $g: \mathsf{T}_Y \rightarrow \mathsf{T}_X$ be a mapping such that:
+	First we reformulate $g$ in terms of the open set.
+We can redefine $g: \mathsf{T}_Y \rightarrow \mathsf{T}_X$ to be the map between the verifiable sets corresponding to the experimental observations.
+This map has the following properties:
 	\begin{enumerate}
@@
 
-	We claim there exists a unique extension $\bar{g}:\sigma(Y)\to\sigma(X)$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma(X)$ and $\sigma(Y)$, such that $\bar{g}|_{\mathsf{T}_Y}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
+	Then we construct the unique extension $\bar{g}:\sigma(Y)\to\sigma(X)$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma(X)$ and $\sigma(Y)$, such that $\bar{g}|_{\mathsf{T}_Y}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
 Let $\bar{g}(V) = g(V)$ for all open sets $V \in \mathsf{T}_Y$.
@@
 
-	We claim we can define $\hat{g}:Y\to\sigma(X)$ such that $\hat{g}(y) = \bar{g}(\{y\})$.
+	Next we define $\hat{g}:Y\to\sigma(X)$ such that $\hat{g}(y) = \bar{g}(\{y\})$.
 Since $Y$ is Hausdorff, every singleton $\{y\}$ is closed and is therefore a Borel set.
-$\bar{g}(\{y\})$ is well defined and so is $\hat{g}(y)$.
+Therefore $\bar{g}(\{y\})$ is well defined and so is $\hat{g}(y)$.
 
-	We claim  $\hat{g}(y_1)\cap\hat{g}(y_2) = \emptyset$ if and only if $y_1\neq y_2$ for all $y_1,y_2\in Y$ such that $\hat{g}(y_i)\neq\emptyset$ for $i=1,2$. If $y_1\neq y_2$ we have
+	We claim that $\hat{g}(y_1)\cap\hat{g}(y_2) = \emptyset$ if and only if $y_1\neq y_2$ for all $y_1,y_2\in Y$ such that $\hat{g}(y_i)\neq\emptyset$ for $i=1,2$. If $y_1\neq y_2$ we have
 	$$
@@
 
-	We claim we can define $f: X\to Y$ such that $f(x) = y$ if and only if $x\in \hat{g}(y)$.
+	We are now ready to define $f: X\to Y$ such that $f(x) = y$ if and only if $x\in \hat{g}(y)$.
 Since $g(Y)=X$, there exists $y\in Y$ such that $x\in\hat{g}(y)$.
@@
 
-	We claim $g = f^{-1} |_{\mathsf{T}(Y)}$.
+	Now we show that $g = f^{-1} |_{\mathsf{T}(Y)}$.
 Let $V\in\mathsf{T}(Y)$.
@@
 
-	We claim $f$ is continuous. It is so since $g = f^{-1} |_{\mathsf{T}(Y)}$ takes open sets to open sets. 
+	Lastly we claim $f$ is continuous. It is so since $g = f^{-1} |_{\mathsf{T}(Y)}$ takes open sets to open sets. 
 \end{proof}
@@
 \end{thm}
+
+\begin{proof}
+	As we saw in the previous result, both definitions lead to experimental relationships being fully characterized by a continuous function.
+\end{proof}
 
@@
 \begin{proof}
-	We claim $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is second-countable.
-We first note that the sub-basis $\{V(U_X, U_Y) \, |\,   U_X \in \mathcal{B}_X , U_Y \in \mathcal{B}_Y \}$ is countable since $\mathcal{B}_X$ and $\mathcal{B}_Y$ are countable.
-The collection $\mathcal{B}$ of all finite intersections is still countable, since it is in one-to-one correspondence with the collection of finite subsets of a countable set, which is still countable.
-Therefore $\mathcal{B}$ is a countable basis, which means $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is second-countable.
+	First we show that $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is second-countable.
+We note that the sub-basis $\{V(U_X, U_Y) \, |\,   U_X \in \mathcal{B}_X , U_Y \in \mathcal{B}_Y \}$ is countable since $\mathcal{B}_X$ and $\mathcal{B}_Y$ are countable and so will be the bases which it generates.
+Which means $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is second-countable.
 
-	We claim $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is Hausdorff.
-Let $f,g:X\to Y$ be distinct continuous functions.
+	Next we show that $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is Hausdorff.
+Let $f,g:X\to Y$ be two distinct continuous functions.
 Then for some $x\in X$, we have $f(x)\neq g(x)$.
@@
 		\hline 
-		$T_2$, 2nd ctbl space. & Experimentally distinguishable space, whose points are the possible values and whose open sets represent the experimentally attainable levels of precision \\
-		Open set & Verifiable set. We can verify experimentally that an element is within the set  \\ 
-		Closed set & Refutable set. We can verify experimentally that an element is not in the set \\ 
-		Basis of top. & A collection of verifiable sets such that any verifiable set is determined by the basis sets\\
-		Continuous \newline function &  A function between two sets of experimental distinguishable elements that preserves distinguishability \\
-		Homeomorphism &  A perfect equivalence between experimentally distinguishable spaces. \\
+		Hausdorff, second-countable space & Space of experimentally distinguishable elements, whose points are the possible values and whose open sets represent the experimentally attainable levels of precision.
+\\
+		Open set & Verifiable set. We can verify experimentally that an element is within the set.  \\ 
+		Closed set & Refutable set. We can verify experimentally that an element is not in the set. \\ 
+		Basis & A collection of verifiable sets that can be used to distinguish an element\\
+		Continuous \newline function &  An experimental relationship between two sets of experimental distinguishable elements, which must preserve distinguishability \\
+		Homeomorphism &  A perfect equivalence between spaces of experimentally distinguishable elements. \\
 	\end{tabular} 

### 2017-07-29 Articles/TopologyAndExperimentalDistinguishability/article.tex
Going through the proofs with Christine

@@
 	\item for each $i=1,\ldots,n$ run the test $\mathsf{e}_i$
-	\item if all tests $\mathsf{e}_i$ terminate successfully, terminate the test $\mathsf{e}_\wedge$ successfully.
+	\item if all tests $\mathsf{e}_i$ terminate successfully then terminate successfully
 	\end{enumerate}
@@
 		\item run the test $\mathsf{e}_i$ for $n$ seconds
-		\item if $\mathsf{e}_i$ terminated successfully, terminate $\mathsf{e}_\vee$ successfully
+		\item if $\mathsf{e}_i$ terminated successfully then terminate successfully
 	\end{enumerate}
@@
 	Suppose there exists an $i \in \mathbb{Z}^+$ such that $\mathsf{e}_i$ will terminate successfully.
-Then the above procedure will eventually run that test for time sufficient for it to terminate successfully.
+Then the above procedure will eventually run that test for sufficient time for it to terminate successfully.
 It will do so in finite time as it will have run finitely many tests finitely many times each for a finite amount of time.
-Therefore $\mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ is an experimental test that is successful if and only if at least one statement in $\{\mathsf{s}_i\}_{i=1}^{n}$ is successful.
-So $\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i =\llparenthesis\bigvee\limits_{i=1}^{\infty} \mathsf{s}_i, \mathsf{e}_{\wedge}(\{\mathsf{e}_i\}_{i=1}^{\infty})\rrparenthesis$ is an experimental observation.
+Therefore $\mathsf{e}_\vee(\{\mathsf{e}_i\}_{i=1}^{\infty})$ is an experimental test that is successful if and only if at least one statement in $\{\mathsf{s}_i\}_{i=1}^{\infty}$ is successful.
+So $\bigvee\limits_{i=1}^{\infty} \mathsf{o}_i =\llparenthesis\bigvee\limits_{i=1}^{\infty} \mathsf{s}_i, \mathsf{e}_{\vee}(\{\mathsf{e}_i\}_{i=1}^{\infty})\rrparenthesis$ is an experimental observation.
 \end{proof}
@@
 Let $U_1, U_2, ...
-, U_n \subseteq X$ be n verifiable sets.
+, U_n \subseteq X$ be $n$ verifiable sets.
 For each $U_i$ there exists an experimental observation $\mathsf{o}_i = \llparenthesis x\in U_i, \mathsf{e}_\in(U_i) \rrparenthesis$.
@@
 Therefore the union of a (countable) basis is the verifiable set $X$.
-Further, there exist at least two incompatible experimental observation, corresponding to two disjoint sets, so the empty set is a verifiable set.
+Further, there exist at least two incompatible experimental observations, corresponding to two disjoint sets, so the empty set is a verifiable set.
 Now, \ref{setbehavior} shows that the collection $\mathsf{T}$ is closed under finite intersection and countable union.
-Because $\mathsf{T}$ is determined by an experimental domain of observations, there is a countable basis of operations which translates to a countable basis of open (verifiable) sets, so it is second-countable.
+Because $\mathsf{T}$ is determined by an experimental domain, there is a countable basis of observations which translates to a countable basis of open (verifiable) sets, so it is second-countable.
 To show it is closed under arbitrary union, notice that an arbitrary union may be rewritten as a union of basis elements, which is then a countable union, and so it remains in $\mathsf{T}$.
@@
 It will be successful if and only if $y \in U_Y$.
-Since $y=f(x)$, it is successful if and only if $x \in f^{-1}(U_Y)$: the procedure is the test $\mathsf{e}_{\in}(f^{-1}(U_Y))$.
+Since $y=f(x)$, it is successful if and only if $x \in f^{-1}(U_Y)$: the procedure is the test $\mathsf{e}_{\in}(f^{-1}(U_Y))$ on $x$.
 This means that $o_x = \llparenthesis x \in f^{-1}(U_Y),\mathsf{e}_{\in}(f^{-1}(U_Y))\rrparenthesis$ is an experimental observation and it must be in $\mathcal{D}_X$ since $\mathcal{D}_X$ contains all possible experimental observations of that form.
@@
 An \textbf{experimental relationship} is a map $g : \mathcal{D}_Y \rightarrow \mathcal{D}_X$ such that if $\mathsf{o} \in \mathcal{D}_Y$ is verified then $g(\mathsf{o}) \in \mathcal{D}_X$ must also be verified.
-To be consistent, such relationship must have these properties:
+To be consistent, such a relationship must have these properties:
 	\begin{enumerate}
@@
 	\begin{enumerate}
-		\item It is compatible with union and intersection, i.e. for any subsets $V_1, V_2 \subseteq Y$, we have $g(V_1 \cup V_2)=g(V_1)\cup g(V_2)$ and $g(V_1 \cap V_2)=g(V_1)\cap g(V_2)$
+		\item it is compatible with union and intersection, i.e. for any subsets $V_1, V_2 \subseteq Y$, we have $g(V_1 \cap V_2)=g(V_1)\cap g(V_2)$ and $g(V_1 \cup V_2)=g(V_1)\cup g(V_2)$
 		\item $g(\emptyset) = \emptyset$
@@
 
-	Then we construct the unique extension $\bar{g}:\sigma(Y)\to\sigma(X)$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma(X)$ and $\sigma(Y)$, such that $\bar{g}|_{\mathsf{T}_Y}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
+	Then we construct the unique extension $\bar{g}:\sigma_Y\to\sigma_X$ to the Borel $\sigma$-algebras of $X$ and $Y$, respectively $\sigma_X$ and $\sigma_Y$, such that $\bar{g}|_{\mathsf{T}_Y}=g$ and $\bar{g}$ is compatible with union, intersection and complements.
 Let $\bar{g}(V) = g(V)$ for all open sets $V \in \mathsf{T}_Y$.
-Let $A \in \sigma(Y)$ (not necessarily open) and $A^C$ be its complement.
+Let $A \in \sigma_Y$ (not necessarily open) and $A^C$ be its complement.
 We must have $\bar{g}(A^C) = \bar{g}(A)^C = X\setminus \bar{g}(A)$ for $\bar{g}$ to be compatible with complements.
-Recall that all Borel sets in $\sigma(Y)$ and $\sigma(X)$ may be written as some combination of unions, intersections, and complements of open sets.
-Thus, the construction uniquely determine what $\bar{g}$ should output on any Borel set.
+Recall that all Borel sets in $\sigma_Y$ and $\sigma_X$ may be written as some combination of unions, intersections, and complements of open sets.
+Thus, the construction uniquely determines what $\bar{g}$ should output on any Borel set.
 We need only check that the output is still a Borel set.
-But by definition of $\bar{g}$, the outputs will be given as unions, intersections, and complements of outputs of $g$, which are open sets, and so the image of $\bar{g}$ is contained in $\sigma(X)$.
+But by definition of $\bar{g}$, the outputs will be given as unions, intersections, and complements of outputs of $g$, which are open sets, and so the image of $\bar{g}$ is contained in $\sigma_X$.
 $\bar{g}$ is well defined.
-The function $\bar{g}$ in a sense represents extracting the maximum amount of information possible out of $g$.
 
-	Next we define $\hat{g}:Y\to\sigma(X)$ such that $\hat{g}(y) = \bar{g}(\{y\})$.
+	Next we define $\hat{g}:Y\to\sigma_X$ such that $\hat{g}(y) = \bar{g}(\{y\})$.
 Since $Y$ is Hausdorff, every singleton $\{y\}$ is closed and is therefore a Borel set.
@@
 $f: X\to Y$ is well defined.
-Note that no arbitrary choice where made so far that lead to the construction of $f$, which is therefore determined uniquely by $g$.
+Note that no arbitrary choices were made so far that led to the construction of $f$, which is therefore determined uniquely by $g$.
 
-	Now we show that $g = f^{-1} |_{\mathsf{T}(Y)}$.
-Let $V\in\mathsf{T}(Y)$.
+	Now we show that $g = f^{-1} |_{\mathsf{T}_Y}$.
+Let $V\in\mathsf{T}_Y$.
 We want to show $f^{-1}(V) = g(V)$.
@@
 $f^{-1}(V) \supseteq g(V)$.
-$f^{-1}(V) = g(V)$ for all $V\in\mathsf{T}(Y)$ and therefore $f^{-1}|_{\mathsf{T}(Y)}=g$.
+$f^{-1}(V) = g(V)$ for all $V\in\mathsf{T}_Y$ and therefore $f^{-1}|_{\mathsf{T}_Y}=g$.
 
-	Lastly we claim $f$ is continuous. It is so since $g = f^{-1} |_{\mathsf{T}(Y)}$ takes open sets to open sets. 
+	Lastly we claim $f$ is continuous. It is so since $g = f^{-1} |_{\mathsf{T}_Y}$ takes open sets to open sets. 
 \end{proof}
@@
 \begin{proof}
-	As we saw in the previous result, both definitions lead to experimental relationships being fully characterized by a continuous function.
+	As we saw in the previous results, both definitions lead to experimental relationships being fully characterized by a continuous function.
 \end{proof}
@@
 Let $\mathcal{B}_X$ and $\mathcal{B}_Y$ be two countable bases for $X$ and $Y$ respectively.
-The basis-to-basis topology $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ on $C(X,Y)$ with respect to the basis $\mathcal{B}_X$ and $\mathcal{B}_Y$ is Hausdorff and second-countable.
+The basis-to-basis topology $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ on $C(X,Y)$ with respect to the bases $\mathcal{B}_X$ and $\mathcal{B}_Y$ is Hausdorff and second-countable.
 \end{prop}
@@
 We note that the sub-basis $\{V(U_X, U_Y) \, |\,   U_X \in \mathcal{B}_X , U_Y \in \mathcal{B}_Y \}$ is countable since $\mathcal{B}_X$ and $\mathcal{B}_Y$ are countable and so will be the bases which it generates.
-Which means $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is second-countable.
+This means $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is second-countable.
 
@@
 We may assume again that $U$ is a basis element for the topology on $X$ by shrinking it if necessary.
-Now, let $T_1$ to be the (sub-)basis element for basis-to-basis topology corresponding to $U$ and $V_1$.
+Now, let $T_1$ be the (sub-)basis element for the basis-to-basis topology corresponding to $U$ and $V_1$.
 By construction, $f\in T_1$.
-Similarly, let $T_2$ to be the basis element for the basis-to-basis topology corresponding to $U$ and $V_2$ and containing $g$.
+Similarly, let $T_2$ be the basis element for the basis-to-basis topology corresponding to $U$ and $V_2$ and containing $g$.
 Since $V_1$ and $V_2$ are disjoint, so are $T_1$ and $T_2$.
-$\mathsf{T}(\mathcal{C})$ is Hausdorff.
+$\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is Hausdorff.
 \end{proof}
@@
 
-What emerges from this work is that \emph{the primary application of topology in science is experimental distinguishability}.
-Its role is to track of what can be distinguished through experimentation.
-The importance of continuous functions in science and physics is explained since experimental relationships have to be consistent with what can be experimentally distinguished.
-This provides a strong justification as to why deterministic systems are always characterized by continuous transformations.
-It is not that they happen to be continuous: they must be.
+What emerges from this work is that the primary application of topology in science is experimental distinguishability: its role is to keep track of what can be distinguished through experimentation.
+The importance of continuous functions in science stems from requiring that experimental relationships be consistent with experimental distinguishability.
+Therefore it is not that the deterministic evolution of a physical system happens to be continuous: it must be.
 
@@
 Funding for this work was provided in part by the MCubed program of the University of Michigan.
-The third author is supported by the National Science Foundation Graduate Research Fellowship Program under Grant No.
-DGE\#1256260.
-Any opinions, findings, and conclusions or recommendations expressed in this material are those of the author(s) and do not necessarily reflect the views of the National Science Foundation.
 

### 2017-07-29 Articles/TopologyAndExperimentalDistinguishability/article.tex
Removing unused symbols and other minor edits

@@
-\documentclass[review]{elsarticle}
+\documentclass[preprint]{elsarticle}
 
@@
 \usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}
-\usepackage{MnSymbol,centernot}
-
-% Symbol for "implies" and "not implies"
-\def\imp{\Rightarrow}
-\def\nimp{\nRightarrow}
-
-% Symbol for "compatibility" and "incompatibility"
-\def\comp{\doublefrown}
-\def\ncomp{\ndoublefrown}
 
@@
 \newtheorem{defn}{Definition}[section]
-\newtheorem{exmp}{Example}[section]
+%\newtheorem{exmp}{Example}[section]
 \theoremstyle{remark}
@@
 
-The successful use of mathematical ideas in experimental sciences is long established and celebrated \cite{wigner}.
-Topology is perhaps the most widespread, either directly\cite{rashevsky, chichilnisky} or as a foundation to other tools\cite{nakahara}.
+The successful use of mathematical ideas in experimental sciences is long established and celebrated~\cite{wigner}.
+Topology is perhaps the most widespread, either directly~\cite{rashevsky, chichilnisky} or as a foundation to other tools~\cite{nakahara}.
 This leads one to ask, why is it so successful?

### 2017-08-01 Blog/09-ClassicalUncertaintyRelationship.tex
New blog post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Classical Uncertainty Relationship}

TL;DR .

In previous posts we have looked at information entropy, the amount of yes/no questions you need to identify an element within a distribution, and the fact that Hamiltonian dynamics conserves that. Here we will show one interest result: because of information entropy information cannot be changed, Hamiltonian mechanics cannot take a distribution and squeeze it as much it wants. Since we always have to use the same number of yes/no questions to identify an element, there is a minimum amount of uncertainty always associated to each element.

Intuitively, if you have a distributions $\rho(q,p)$ over phase space and you shrink it, the information will diminish: you will need less information to identify an element. That type of transformation cannot done through Hamiltonian evolution since it conserves information entropy. We are going to ask this question: given a fixed amount $I_0$ of information, what is the distribution $\rho(q,p)$ that minimizes the variance? We will find that it is the Gaussian distribution where the product $\sigma_q\sigma_p$ of the standard deviations to the entropy is given by the following relationship:

\begin{equation}
\sigma_x\sigma_p \geq \exp (I_0 - 1) / 2 \pi 
\label{classicalUncertainty}
\end{equation}

Note the similarity with the Heisenberg uncertainty principle of quantum mechanics:

\begin{equation}
\sigma_x\sigma_p \geq \hbar / 2
\label{quantumUncertainty}
\end{equation}

\section{Minimizing uncertainty}

Suppose we have a phase space distribution $\rho(x,p)$ As the state evolves into its final state, the amount of material remains the same: $\int \rho \, dx \wedge dp = 1$. The final entropy will also remain the same: $I(\rho) = \int \rho \ln \rho \, dx \wedge dp = I_0$. We now ask the following question: once we fix the entropy, what is the distribution that minimizes the product $\sigma_x\sigma_p$ of the standard deviations? That would correspond to the minimum uncertainty for the final distribution given by Hamiltonian evolution on the initial distribution.

We are going to use the Lagrange multipliers, which is a standard technique in calculus of variations. We write a function that is the product of the variances $\sigma_x^2 \sigma_p^2 \equiv \int (x-\mu_x)^2 \rho \, dxdp \int (p-\mu_p)^2 \rho \, dxdp$ summed with the two constraints (the total amount of material is $1$ and the total entropy $I_0$) each multiplied by a constant. Then take the derivative and set it to $0$.
\begin{align*}
L = &\int (x-\mu_x)^2 \rho \, dxdp \int (p-\mu_p)^2 \rho \, dxdp \\
&+ \lambda_1\left(\int \rho dxdp - 1\right) + \lambda_2\left(- \int \rho \ln \rho \, dxdp - I_0\right)\\
\delta L = &\int \delta \rho [(x-\mu_x)^2 \sigma_p^2 + \sigma_x^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho - \lambda_2 ] dxdp = 0 \\
\lambda_2 \ln \rho = &\lambda_1 - \lambda_2 + (x-\mu_x)^2 \sigma_p^2 + \sigma_x^2 (p-\mu_p)^2 \\
\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(x-\mu_x)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_x^2 (p-\mu_p)^2}{\lambda_2}}\\
\end{align*}
Now that we found the form of the function, we solve the multipliers. That is: we find $\lambda_1$ and $\lambda_2$ such that the distribution is unitary and the information entropy is $I_0$. We have:
\begin{align*}
\rho = &\frac{1}{ 2 \pi \sigma_x \sigma_p} e^{-\frac{(x-\mu_x)^2}{2\sigma_x^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
I_0 = &\ln (2\pi\sigma_x\sigma_p) + 1
\end{align*}
We find that the distribution that minimizes the spread is the product of two independent Gaussians. Recall that in quantum mechanics, the Gaussian wave packet is the state that minimizes uncertainty: this is the classical analogue.

As the distribution evolves, the entropy is conserved and therefore the product $\sigma_x^2 \sigma_p^2$ can never be less than the one given by the Gaussian distribution of the same entropy. We have:
\begin{align*}
\sigma_x\sigma_p \geq \exp (I_0 - 1) / 2 \pi 
\end{align*}
This is strikingly similar to the Heisenberg uncertainty principle, except that the value depends on the initial entropy $I_0$.

\section{Conclusion}

We have found that also classical Hamiltonian mechanics satisfies an uncertainty relationship. Is this a coincidence? Or is there something more?

This is my take: the difference between the classical case and the quantum case is that the uncertainty is set by the initial conditions in the first while it is universal in the second. Now, suppose that each element of the classical distribution was subjected to a universal but totally chaotic force. Then we would not be able to create initial conditions with an entropy lower than the chaotic force would allow. For the same token, we also would not be able to assign an infinitely precise position and momentum to each element of the distribution given that it keeps jiggling due to the chaotic force.

If the entropy associated to the initial distribution is large compared to the entropy associated with chaotic force, then we can disregard it and classical mechanics hold. If not, then the classical description starts to break down.

In this picture, there is nothing weird going on. Classical and quantum mechanics are very related to each other, both describing the evolution of distributions that conserve information entropy.

\end{document}

### 2017-08-01 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
More work on HamMech article

@@
 
+The reader may have notice that there the proof of the single degree of freedom was trivial: just a mere application of the definitions.
+The clever bit was to set up the problem.
+In the multiple degrees of freedom setting up the problem presents a challenge: we don't have a suitable mathematical tool to do it.
+
+Mathematically, Hamiltonian mechanics is formalized with symplectic geometry, a branch of differential geometry.
+But while symplectic geometry gives us the right tool to study the geometry of the volumes and areas on which we define the distributions, it does not give us good tools to study the distributions themselves.
+This because differential geometry in general focuses on objects that are coordinate independent while distributions are coordinate dependent.
+On the other hand, probability theory and statistics give us the right tool to study distributions on multiple degrees of freedom but they do not give us tools to define areas and volumes.
+So, to make a mathematically precise proof, we would need to combine elements of both and once the problem is correctly setup, the proof would be trivial.
+Unfortunately, the result would be something difficult to follow and would make the physical content opaque.
+
+What we are going to do, instead, is to show how and why the math works the way that it does and what the key intuitive ideas areas.
+It will also give enough elements for someone to construct a more formal proof, in case it is needed.
+
 Show conceptually what needs to happen.

### 2017-08-07 Articles/TopologyAndExperimentalDistinguishability/article.tex
Added reference to Basener textbook
Removing "Submitted" footnote
Addressing observation from Mayer

@@
 \documentclass[preprint]{elsarticle}
+
+% Temporary to remove "submitted to"
+\makeatletter
+\def\ps@pprintTitle{%
+	\let\@oddhead\@empty
+	\let\@evenhead\@empty
+	\def\@oddfoot{}%
+	\let\@evenfoot\@oddfoot}
+\makeatother
 
@@
 The successful use of mathematical ideas in experimental sciences is long established and celebrated~\cite{wigner}.
-Topology is perhaps the most widespread, either directly~\cite{rashevsky, chichilnisky} or as a foundation to other tools~\cite{nakahara}.
+Topology is perhaps the most widespread, either directly~\cite{rashevsky, chichilnisky} or as a foundation to other tools~\cite{nakahara, basener}.
 This leads one to ask, why is it so successful?
@@
 
+\begin{proof}
+	We need to show that the basis-to-basis topology is a topology.
+By definition, it is closed under finite intersection and arbitrary union.
+The set $V(X,\emptyset)$ is in the topology and contains no functions.
+The set $V(X,Y)$ is in the topology and contains all functions.
+\end{proof}
+
 \begin{prop}
@@
 
-We hope that this work, because of both its method and result, can provide a more solid foundation to formalize experimental sciences.
+In light of this work, the effectiveness of mathematics in the natural sciences is perhaps not so unreasonable.
+We hope that the methods and results shown here can provide a more solid foundation to formalize experimental sciences.
 

### 2017-08-07 Articles/TopologyAndExperimentalDistinguishability/article.tex
Using Mark's wording

@@
 \begin{proof}
-	We need to show that the basis-to-basis topology is a topology.
-By definition, it is closed under finite intersection and arbitrary union.
-The set $V(X,\emptyset)$ is in the topology and contains no functions.
-The set $V(X,Y)$ is in the topology and contains all functions.
+	The collection $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is defined to be the topology generated by the sets $V(U_X,U_Y)$, so it contains the empty set, all of $C(X,Y)$, and is closed under arbitrary union and finite intersection by definition.
 \end{proof}

### 2017-08-08 Articles/TopologyAndExperimentalDistinguishability/article.tex
Finished proof for basis-to-basis topology

@@
 \begin{proof}
-	The collection $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is defined to be the topology generated by the sets $V(U_X,U_Y)$, so it contains the empty set, all of $C(X,Y)$, and is closed under arbitrary union and finite intersection by definition.
+	The collection $\mathsf{T}(C(X,Y), \mathcal{B}_X, \mathcal{B}_Y)$ is defined to be the topology generated by the sets $V(U_X,U_Y)$, so it contains the empty set and is closed under arbitrary union and finite intersection by definition.
+To see why these sets contain every continuous function, let $f\in C(X,Y)$.
+Then for any $U_Y\in \mathcal{B}_Y$, we can find some $U_X\in \mathcal{B}_X$ such that $U_X\subseteq f^{-1}(U_Y)$.
+Then $f\in V(U_X,U_Y)$.
 \end{proof}

### 2017-08-14 Blog/09-ClassicalUncertaintyRelationship.tex
Reviewed blog post

@@
 
-TL;DR .
+TL;DR Classical Hamiltonian mechanics already includes an uncertainty relationship that is similar to Heisenberg's uncertainty principle of quantum mechanics.
 
-In previous posts we have looked at information entropy, the amount of yes/no questions you need to identify an element within a distribution, and the fact that Hamiltonian dynamics conserves that.
-Here we will show one interest result: because of information entropy information cannot be changed, Hamiltonian mechanics cannot take a distribution and squeeze it as much it wants.
+In previous posts we have looked at information entropy, the number of yes/no questions you need to identify an element within a distribution, and the fact that Hamiltonian dynamics conserves that.
+Here we will show one interesting result: since information entropy cannot be changed, Hamiltonian mechanics cannot take a distribution and squeeze it arbitrarily.
 Since we always have to use the same number of yes/no questions to identify an element, there is a minimum amount of uncertainty always associated to each element.
 
-Intuitively, if you have a distributions $\rho(q,p)$ over phase space and you shrink it, the information will diminish: you will need less information to identify an element.
-That type of transformation cannot done through Hamiltonian evolution since it conserves information entropy.
-We are going to ask this question: given a fixed amount $I_0$ of information, what is the distribution $\rho(q,p)$ that minimizes the variance?
-We will find that it is the Gaussian distribution where the product $\sigma_q\sigma_p$ of the standard deviations to the entropy is given by the following relationship:
+Intuitively, if you have a distribution $\rho(x,p)$ over phase space and you shrink it, the information will diminish: you will need less information to identify an element.
+That type of transformation cannot be done through Hamiltonian evolution since it conserves information entropy.
+We are going to ask this question: given a fixed amount $I_0$ of information entropy, what is the distribution $\rho(x,p)$ that minimizes the variance?
+We will find that it is the Gaussian distribution, and therefore in general the product $\sigma_x\sigma_p$ of the standard deviations is related to the information entropy by the following relationship:
 
@@
 
+Let's see how it works.
+
 \section{Minimizing uncertainty}
 
-Suppose we have a phase space distribution $\rho(x,p)$ As the state evolves into its final state, the amount of material remains the same: $\int \rho \, dx \wedge dp = 1$.
+Suppose we have a phase space distribution $\rho(x,p)$.
+As the state evolves into its final state, the amount of material remains the same: $\int \rho \, dx \wedge dp = 1$.
 The final entropy will also remain the same: $I(\rho) = \int \rho \ln \rho \, dx \wedge dp = I_0$.
@@
 
-We are going to use the Lagrange multipliers, which is a standard technique in calculus of variations.
+We are going to use Lagrange multipliers, which is a standard technique in calculus of variations.
 We write a function that is the product of the variances $\sigma_x^2 \sigma_p^2 \equiv \int (x-\mu_x)^2 \rho \, dxdp \int (p-\mu_p)^2 \rho \, dxdp$ summed with the two constraints (the total amount of material is $1$ and the total entropy $I_0$) each multiplied by a constant.
@@
 \begin{align*}
+L = &\sigma_x^2 \sigma_p^2 + \lambda_1\left(\int \rho dxdp - 1\right) + \lambda_2\left(- \int \rho \ln \rho \, dxdp - I_0\right)\\
+\delta L = & \delta \sigma_x^2 \sigma_p^2 + \sigma_x^2 \delta \sigma_p^2 +  \lambda_1 \, \delta \int \rho dxdp - \lambda_2 \, \delta \int \rho \ln \rho \, dxdp \\
+= &\int  (x-\mu_x)^2 \delta \rho \, dxdp \, \sigma_p^2 +  \sigma_x^2 \int (p-\mu_p)^2 \delta \rho \, dxdp + \\ &\lambda_1 \int \delta\rho \, dxdp - \lambda_2 \int (\delta \rho \ln \rho + \rho \delta \ln \rho)dxdp  \\
+= &\int [(x-\mu_x)^2 \sigma_p^2 + \sigma_x^2 (p-\mu_p)^2 + \lambda_1 - \lambda_2 \ln \rho - \lambda_2 ] \delta \rho  \, dxdp = 0 \\
+\lambda_2 \ln \rho = &\lambda_1 - \lambda_2 + (x-\mu_x)^2 \sigma_p^2 + \sigma_x^2 (p-\mu_p)^2 \\
+\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(x-\mu_x)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_x^2 (p-\mu_p)^2}{\lambda_2}}\\
 L = &\int (x-\mu_x)^2 \rho \, dxdp \int (p-\mu_p)^2 \rho \, dxdp \\
 &+ \lambda_1\left(\int \rho dxdp - 1\right) + \lambda_2\left(- \int \rho \ln \rho \, dxdp - I_0\right)\\
-\delta L = &\int \delta \rho [(x-\mu_x)^2 \sigma_p^2 + \sigma_x^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho - \lambda_2 ] dxdp = 0 \\
-\lambda_2 \ln \rho = &\lambda_1 - \lambda_2 + (x-\mu_x)^2 \sigma_p^2 + \sigma_x^2 (p-\mu_p)^2 \\
-\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(x-\mu_x)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_x^2 (p-\mu_p)^2}{\lambda_2}}\\
 \end{align*}
@@
 Then we would not be able to create initial conditions with an entropy lower than the chaotic force would allow.
-For the same token, we also would not be able to assign an infinitely precise position and momentum to each element of the distribution given that it keeps jiggling due to the chaotic force.
+For the same reason, we also would not be able to track infinitely precisely position and momentum of each element of the distribution given that it keeps jiggling due to the chaotic force.
 
-If the entropy associated to the initial distribution is large compared to the entropy associated with chaotic force, then we can disregard it and classical mechanics hold.
+If the entropy associated to the initial distribution is large compared to the entropy associated with chaotic force, then we can disregard it and classical mechanics holds.
 If not, then the classical description starts to break down.
 
-In this picture, there is nothing weird going on. Classical and quantum mechanics are very related to each other, both describing the evolution of distributions that conserve information entropy.
+In this picture, classical and quantum mechanics are not so different after all: they both describe the evolution of distributions that conserve information entropy.
 

### 2017-08-14 Blog/10-GalileanHamiltonianMechanicsDoesntReallyExist.tex
Post on Galilean boost and Hamiltonian mechanics

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Galilean Hamiltonian mechanics doesn't really exist}

or "Non-relativistic Hamiltonian mechanics doesn't really exist"
or "All Hamiltonian mechanics is relativistic"

TL;DR Galilean transformations are not compatible with standard Hamiltonian mechanics. They give good approximations only for small changes in velocity.

As part of my work in foundations of physics, I am forced to constantly search for different ways to look at things. Sometimes this leads me to have a different perspective which brings a more consistent overall picture but forces one to reshuffle some of the basic concepts. The relationship between relativity and Hamiltonian mechanics is one of those cases.

When studying physics, you learn that relativistic mechanics is more ``correct" and non-relativistic mechanics is just an approximation. But you may also get a sense that non-relativistic mechanics is self-consistent: it just so happens that the world is different. If you look at things closely, though, I don't think that's the case.

First we have to make a distinction between the ``principle of relativity", which just states that the laws are invariant under any coordinate transformations, and the ``theory of special relativity", which adds that the speed of light is invariant and gives you things like $E=mc^2$. We are mainly going to concentrate on the first.

The idea is this: if we make Hamiltonian mechanics ``relativistic", in the sense of making sure it gives the same equations for all reference frames, what do we get? It will take a few posts to explore this question. What we'll see is that we already get some results associated with the theory of special relativity, even though we didn't require the invariance of the speed of light.

We start by concentrating on a simpler question: is standard Hamiltonian mechanics invariant under Galilean transformations? That is, what happens if we boost our frame according to $\hat{x}=x+ v_0 t$? What we find is that standard Hamiltonian mechanics already breaks down and the transformation is a good approximation only if $v_0$ is small. Which is a result that already sounds like the theory of special relativity.

Let's break this down into smaller steps.

\section{Change of coordinates}

Suppose we have a Hamiltonian system. Its state is given by position $x$ and momentum $p$. Its evolution is given by:
\begin{equation}
\begin{aligned}
\frac{dx^i}{dt} = \frac{\partial H}{\partial p_i}  \\
\frac{dp_i}{dt} = - \frac{\partial H}{\partial x^i}  \\
\end{aligned}
\label{Hamilton}
\end{equation}
where $H(x^i,p_j)$ is the Hamiltonian.

Suppose we change coordinates $\hat{x}^i = \hat{x}^i(x^j)$: how does the momentum change? We can recover the transformation rule of momentum by requiring that, in the new coordinates, the system is still Hamiltonian. One way of doing that is requiring that area elements of phase space are conserved. We find that the product of two matrices $\frac{\partial \hat{x}^i}{\partial x^j}$ and $\frac{\partial \hat{p}_i}{\partial p_k}$ must be the identity matrix, which means one is the inverse of the other.
\begin{equation*}
\begin{aligned}
dx^i \wedge dp_i = d\hat{x}^i\wedge d\hat{p}_i &=  \frac{\partial \hat{x}^i}{\partial x^j} dx^j \wedge \frac{\partial \hat{p}_i}{\partial p_k} dp_k \\
\frac{\partial \hat{x}^i}{\partial x^j} \frac{\partial \hat{p}_i}{\partial p_k} &= \delta^k_j \\
\frac{\partial \hat{p}_i}{\partial p_k} = \left[\frac{\partial \hat{x}^i}{\partial x^k}\right]^{-1} &=  \frac{\partial x^k}{\partial \hat{x}^i}
\end{aligned}
\end{equation*}

This means that $\frac{\partial \hat{p}_i}{\partial p_k}$ is only a function of position, therefore the new momentum must be a linear transformation. Integrating we have:
\begin{equation*}
\begin{aligned}
\hat{x}^i &= \hat{x}^i(x^j) \\
\hat{p}^i &= \frac{\partial x^j}{\partial \hat{x}^i} p_j + \frac{\partial f(x^j)}{\partial \hat{x}^i}
\end{aligned}
\end{equation*}
where $f$ is an arbitrary function. Since it is arbitrary, we set it to zero and have:
\begin{equation}
\begin{aligned}
\hat{x}^i &= \hat{x}^i(x^j) \\
\hat{p}^i &= \frac{\partial x^j}{\partial \hat{x}^i} p_j
\end{aligned}
\label{spaceTransf}
\end{equation}
Technically, we haven't shown why the arbitrary function added to the momentum is actually the gradient of a scalar function. Since we are setting it to zero, I didn't want to put details that are not that relevant.

The point is: the components of momentum do not transform like a vector, they are not multiplied by the Jacobian, but do transform like a covector, they are multiplied by the inverse of the Jacobian. $dx^i$ gets multiplied by the Jacobian, $dp_i$ by its inverse, so the phase space area remains the same. Note that the transformation rules are completely general: they do not depend on the particular $\hat{x}^i(x^j)$ or the particular $H(x^i,p_j)$. They are valid across all coordinate frames for any Hamiltonian system. As far as spatial transformation is concerned, Hamiltonian mechanics is ``relativistic" as it respects the principle of relativity for those transformations. Note that the same can't be said for Newtonian mechanics.

\section{Galilean transformations}
Now we want to see what happens if we use a Galilean boost and, for simplicity, we'll only use the coordinate in the direction of boost. Plugging into \eqref{spaceTransf} we have:
\begin{equation}
\begin{aligned}
\hat{x} &= x + v_0 t \\
\hat{p} &= p
\end{aligned}
\label{galTransf}
\end{equation}
since $\frac{\partial x}{\partial \hat{x}} = \frac{\partial (\hat{x} - v_0 t)}{\partial \hat{x}} = 1$. Already note something unexpected: momentum does not change. We could use the arbitrary function and set $\hat{p} = p + m v_0$ but that is really ad-hoc, it's not following the general rule.

If we want to calculate the differential and the derivative, we have to decide how to treat $t$. We could treat it as a constant parameter and have:
\begin{equation}
\begin{aligned}
d\hat{x} &= dx \\
\frac{d\hat{x}}{dt} &= \frac{dx}{dt} \\
d\hat{x}\wedge d\hat{p} &= dx \wedge dp
\end{aligned}
\label{timeParameter}
\end{equation}
Or we can treat is as an independent variable:
\begin{equation}
\begin{aligned}
d\hat{x} &= dx + v_0 dt \\
\frac{d\hat{x}}{dt} &= \frac{dx}{dt} + v_0 \\
d\hat{x}\wedge d\hat{p} &= dx \wedge dp + v_0 dt \wedge dp
\end{aligned}
\label{timeParameter}
\end{equation}

In both cases we have a problem. In the first, the velocity does not change as we are adding the same constant number throughout the trajectory. In the second, the area of phase space is not conserved, so the coordinate transformation is not canonical: the equations of motion are not going to be Hamilton's equations so they are not invariant.

Note, though, that if $v_0$ is ``small enough", the area of phase space is ``almost" conserved so the error introduced by using Hamilton's equations is ``small". In other words: the transformation is still incorrect but not \emph{that} incorrect. Standard Hamiltonian mechanics with Galilean transformation can be a good approximation. But if $v_0$ is large, it breaks down.

\section{Conclusion}

If you tried with more complicated transformations, you'd get even worse results. In the sense explored here, standard Hamiltonian mechanics is not relativistic under any time transformation, including Galilean. It's not that we need to make Hamiltonian mechanics relativistic so that it can handle the invariance of the speed of light: we need to make it relativistic so that it can handle any type of coordinate transformation. The invariance of the speed of light does something different: it constrains the Hamiltonian.

This issue is not explored in textbooks. But they do change coordinate systems, even to non-inertial ones! How do they do it? What they do is start from the Lagrangian, which is invariant under all transformations, change the variable there, calculate the new conjugate momentum and then the new Hamiltonian. This does indeed yield the correct result, so if you are only interested in crunching the numbers you are done. That is why, at a technical level, they always put the Lagrangian first. But it does have at least two problems.

First, as we saw in other posts, there are Hamiltonian systems for which you can't write the Lagrangian. In practice, all interesting systems are also Lagrangian so this is more of a conceptual problem. But conceptual problems are exactly what we need to solve if we want to get a good understanding.

The second problem is that it hides the issue, which prevents getting a clear picture. If one addresses it head on, it gives better insights on the connections between classical mechanics, quantum mechanics and special relativity. And that is what we are going to see in the next few posts.

\end{document}

### 2017-08-18 Blog/09-ClassicalUncertaintyRelationship.tex
Reviewed blog posts

@@
 As the state evolves into its final state, the amount of material remains the same: $\int \rho \, dx \wedge dp = 1$.
-The final entropy will also remain the same: $I(\rho) = \int \rho \ln \rho \, dx \wedge dp = I_0$.
+The final entropy will also remain the same: $I(\rho) = - \int \rho \ln \rho \, dx \wedge dp = I_0$.
 We now ask the following question: once we fix the entropy, what is the distribution that minimizes the product $\sigma_x\sigma_p$ of the standard deviations?
@@
 We are going to use Lagrange multipliers, which is a standard technique in calculus of variations.
-We write a function that is the product of the variances $\sigma_x^2 \sigma_p^2 \equiv \int (x-\mu_x)^2 \rho \, dxdp \int (p-\mu_p)^2 \rho \, dxdp$ summed with the two constraints (the total amount of material is $1$ and the total entropy $I_0$) each multiplied by a constant.
+We write a function $M$ that is the product of the variances $\sigma_x^2 \sigma_p^2 \equiv \int (x-\mu_x)^2 \rho \, dxdp \int (p-\mu_p)^2 \rho \, dxdp$ summed with the two constraints (the total amount of material is $1$ and the total entropy $I_0$) each multiplied by a constant.
 Then take the derivative and set it to $0$.
 \begin{align*}
-L = &\sigma_x^2 \sigma_p^2 + \lambda_1\left(\int \rho dxdp - 1\right) + \lambda_2\left(- \int \rho \ln \rho \, dxdp - I_0\right)\\
-\delta L = & \delta \sigma_x^2 \sigma_p^2 + \sigma_x^2 \delta \sigma_p^2 +  \lambda_1 \, \delta \int \rho dxdp - \lambda_2 \, \delta \int \rho \ln \rho \, dxdp \\
+M = &\sigma_x^2 \sigma_p^2 + \lambda_1\left(\int \rho dxdp - 1\right) + \lambda_2\left(- \int \rho \ln \rho \, dxdp - I_0\right)\\
+\delta M = & \delta \sigma_x^2 \sigma_p^2 + \sigma_x^2 \delta \sigma_p^2 +  \lambda_1 \, \delta \int \rho dxdp - \lambda_2 \, \delta \int \rho \ln \rho \, dxdp \\
 = &\int  (x-\mu_x)^2 \delta \rho \, dxdp \, \sigma_p^2 +  \sigma_x^2 \int (p-\mu_p)^2 \delta \rho \, dxdp + \\ &\lambda_1 \int \delta\rho \, dxdp - \lambda_2 \int (\delta \rho \ln \rho + \rho \delta \ln \rho)dxdp  \\
@@
 \lambda_2 \ln \rho = &\lambda_1 - \lambda_2 + (x-\mu_x)^2 \sigma_p^2 + \sigma_x^2 (p-\mu_p)^2 \\
-\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(x-\mu_x)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_x^2 (p-\mu_p)^2}{\lambda_2}}\\
-L = &\int (x-\mu_x)^2 \rho \, dxdp \int (p-\mu_p)^2 \rho \, dxdp \\
-&+ \lambda_1\left(\int \rho dxdp - 1\right) + \lambda_2\left(- \int \rho \ln \rho \, dxdp - I_0\right)\\
+\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(x-\mu_x)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_x^2 (p-\mu_p)^2}{\lambda_2}}
 \end{align*}

### 2017-08-18 Blog/10-GalileanHamiltonianMechanicsDoesntReallyExist.tex
Reviewed blog posts

@@
 
-\section{Galilean Hamiltonian mechanics doesn't really exist}
+\section{Galilean transformations are not compatible with Hamiltonian mechanics}
 
-or "Non-relativistic Hamiltonian mechanics doesn't really exist"
-or "All Hamiltonian mechanics is relativistic"
-
-TL;DR Galilean transformations are not compatible with standard Hamiltonian mechanics. They give good approximations only for small changes in velocity.
-
-As part of my work in foundations of physics, I am forced to constantly search for different ways to look at things.
-Sometimes this leads me to have a different perspective which brings a more consistent overall picture but forces one to reshuffle some of the basic concepts.
-The relationship between relativity and Hamiltonian mechanics is one of those cases.
+TL;DR Galilean boosts do not leave Hamilton's equations and phase space volumes unchanged. They give good approximations only for small changes in velocity.
 
@@
 If you look at things closely, though, I don't think that's the case.
+The problem is that typically in textbooks a small sleight of hand is performed, so we have to look from a slightly different angle to see the trick.
 
@@
 
-Suppose we have a Hamiltonian system. Its state is given by position $x$ and momentum $p$. Its evolution is given by:
+Suppose we have a Hamiltonian system. Its state is given by position $x^i$ and momentum $p_j$. Its evolution is given by:
 \begin{equation}
@@
 \frac{dx^i}{dt} = \frac{\partial H}{\partial p_i}  \\
-\frac{dp_i}{dt} = - \frac{\partial H}{\partial x^i}  \\
+\frac{dp_j}{dt} = - \frac{\partial H}{\partial x^j}  \\
 \end{aligned}
@@
 One way of doing that is requiring that area elements of phase space are conserved.
-We find that the product of two matrices $\frac{\partial \hat{x}^i}{\partial x^j}$ and $\frac{\partial \hat{p}_i}{\partial p_k}$ must be the identity matrix, which means one is the inverse of the other.
+We find that the product of the two matrices $\frac{\partial \hat{x}^i}{\partial x^j}$ and $\frac{\partial \hat{p}_i}{\partial p_k}$ must be the identity matrix, which means one is the inverse of the other.
 \begin{equation*}
@@
 
-This means that $\frac{\partial \hat{p}_i}{\partial p_k}$ is only a function of position, therefore the new momentum must be a linear transformation. Integrating we have:
+This means that $\frac{\partial \hat{p}_i}{\partial p_k}$ is only a function of position, therefore the new momentum must be a linear transformation with respect to the old momentum.
+Integrating we have:
 \begin{equation*}
@@
 \hat{x}^i &= \hat{x}^i(x^j) \\
-\hat{p}^i &= \frac{\partial x^j}{\partial \hat{x}^i} p_j + \frac{\partial f(x^j)}{\partial \hat{x}^i}
+\hat{p}_i &= \frac{\partial x^j}{\partial \hat{x}^i} p_j + \frac{\partial f(x^j)}{\partial \hat{x}^i}
 \end{aligned}
@@
 \hat{x}^i &= \hat{x}^i(x^j) \\
-\hat{p}^i &= \frac{\partial x^j}{\partial \hat{x}^i} p_j
+\hat{p}_i &= \frac{\partial x^j}{\partial \hat{x}^i} p_j
 \end{aligned}

### 2017-08-18 Articles/TopologyAndExperimentalDistinguishability/article.tex
Final formatting changes
Fixing numbering of mathematical statements
Version to submit to arXiv

@@
 \newtheorem{prop}[thm]{Proposition} 
-\newtheorem{cor}{Corollary}
+\newtheorem{cor}[thm]{Corollary}
 
 \theoremstyle{definition}
-\newtheorem{defn}{Definition}[section]
+\newtheorem{defn}[thm]{Definition}
 %\newtheorem{exmp}{Example}[section]
@@
 
-\author[]{Christine A. Aidala\corref{cor1}}
+\author[]{Christine A. Aidala}
 \ead{caidala@umich.edu}
 
-\author[]{Gabriele Carcassi}
+\author[]{Gabriele Carcassi\corref{cor1}}
 \ead{carcassi@umich.edu}
@@
 
-\linenumbers
+%\linenumbers
 
@@
 The notion of ``nearness" captured by topologies keeps track of how hard it is to tell two elements apart via scientific observation.
-The standard topology on a real line, for example, captures the inability to measure a continuous value with infinite precision.
+The standard topology on the real line, for example, captures the inability to measure a continuous value with infinite precision.
 This connection holds in cases where a metric would not be physically meaningful (e.g.~phase space) and in discrete spaces (where each element can be individually identified).
@@
 We then study their properties under logical operations and conclude that they are only closed under finite conjunction and countable disjunction.
-Next we define experimental distinguishability and show, as our fist main result, that any set of objects that are experimentally distinguishable is a Hausdorff, second-countable topological space, where the topology is formally defined in terms of the observations themselves.
+Next we define experimental distinguishability and show, as our first main result, that any set of objects that are experimentally distinguishable is a Hausdorff, second-countable topological space, where the topology is formally defined in terms of the observations themselves.
 Last we define experimental relationships between experimentally distinguishable objects and show, as our second main result, that experimental relationships are represented by continuous functions and are themselves experimentally distinguishable.
@@
 \begin{defn}
-	An \textbf{experimental identification domain} is the triplet $(\mathcal{D}_X, X, x)$ where:
+	An \textbf{experimental identification domain} is the triplet \linebreak $(\mathcal{D}_X, X, x)$ where:
 	\begin{itemize}
@@
 \begin{defn}[Experimental relationship between possibilities]
-	Let $(\mathcal{D}_X, X, x)$ and $(\mathcal{D}_X, Y, y)$ be two experimental identification domains.
+	Let \linebreak $(\mathcal{D}_X, X, x)$ and $(\mathcal{D}_Y, Y, y)$ be two experimental identification domains.
 An \textbf{experimental relationship} is a map $f : X \rightarrow Y$ that can be used within an experimental test.
@@
 $f: X\to Y$ is well defined.
-Note that no arbitrary choices were made so far that led to the construction of $f$, which is therefore determined uniquely by $g$.
+Note that no arbitrary choices were made that led to the construction of $f$, which is therefore determined uniquely by $g$.
 
@@
 		Basis & A collection of verifiable sets that can be used to distinguish an element\\
-		Continuous \newline function &  An experimental relationship between two sets of experimental distinguishable elements, which must preserve distinguishability \\
+		Continuous \newline function &  An experimental relationship between two sets of experimentally distinguishable elements, which must preserve distinguishability \\
 		Homeomorphism &  A perfect equivalence between spaces of experimentally distinguishable elements. \\
 	\end{tabular} 
-	\caption{Topology to physics dictionary. This table sums up the relationship established between mathematical and scientific concepts.}
+	\caption{Topology-to-physics dictionary. This table sums up the relationships established between mathematical and scientific concepts.}
 \end{table}

### 2017-09-11 Blog/11-HamiltoniansAndCoordinateChanges.tex
New blog post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Hamiltonians and coordinate changes}

TL;DR Hamiltonians are not invariant: they change as the time component of a covector (i.e. covariant component) in phase space.

In the previous post we saw how momentum changes as covariant components and that keeps the Hamiltonian equations unchanged under coordinate transformations. We have also seen, though, that under coordinate transformation that mix time and space we have a problem.

What we'll show in this post is that if we impose that Hamilton's equations remain unchanged for all coordinate transformations then the Hamiltonian itself is not invariant but it changes as the covariant time component of a covector. That is, in the same way we group space and time $(x, t)$, we also group momentum and Hamiltonian into a covector of components $(p, -H)$.

Let's see how this work.

\section{Change of space-time coordinates}

Unfortunately, I haven't found a way to derive the transformation laws that only uses simple vector calculus. At least, not one that generalizes well. Let's at least intuitively try to understand why the Hamiltonian cannot possibly be invariant.

Recall Hamilton's equations:
\begin{equation}
\begin{aligned}
\frac{dx^i}{dt} = \frac{\partial H}{\partial p_i}  \\
\frac{dp_j}{dt} = - \frac{\partial H}{\partial x^j}  \\
\end{aligned}
\label{Hamilton}
\end{equation}
where $H(x^i,p_j)$ is the Hamiltonian. Note how space and momentum appear on each side of each equation. Note, though, that they always appear inverted: when one is in the numerator, the other is in the denominator and vice-versa. That is way they have to change in the opposite way: if space is contravariant while momentum is covariant then the equations stay the same.

If we change time (e.g. imagine changing units by multiplying by a constant) and nothing else changes than the equations change as well. Note, though, that the Hamiltonian appears inverted on the opposite side of the equation: time is always in the denominator while the Hamiltonian is always in the numerator. Intuitively: if we want the equations not to change, then the Hamiltonian must change in a way opposite of the time change.

In a previous post, we also saw how Hamilton's equations for a single degree of freedom can also be written as:
\begin{equation}
\begin{aligned}
\vec{S} &= \dfrac{d}{dt} [x, p, t] =  - \mathrm{curl}(\vec{\theta}) = - \mathrm{curl}([p, 0, -H]) \\
S^x &= \dfrac{dx}{dt} = - \partial_p \theta_t + \partial_t \theta_p = \partial_p H + \partial_t 0 = \partial_p H \\
S^p &= \dfrac{dp}{dt} = - \partial_t \theta_x + \partial_x \theta_t = -\partial_t p - \partial_x H = - \partial_x H \\
S^t &= \dfrac{dt}{dt} = - \partial_x \theta_p + \partial_p \theta_x = - \partial_x 0 + \partial_p p = 1
\end{aligned}
\label{HamiltonEquations}
\end{equation}
The covector $\vec{\theta}=[p, 0, -H]$ functions as a vector potential for the displacement vector field $\vec{S} = \dfrac{d}{dt} [x, p, t]$. So, again, we see that the Hamiltonian has to change as the covariant time component. Since $\vec{\theta}$ is a covector, the components change with the inverse of the Jacobian:
\begin{equation}
\begin{aligned}
\theta_{\hat{x}} &= \frac{\partial x}{\partial \hat{x}}\theta_x + \frac{\partial p}{\partial \hat{x}}\theta_p + \frac{\partial t}{\partial \hat{x}}\theta_t \\
\theta_{\hat{p}} &= \frac{\partial x}{\partial \hat{p}}\theta_x + \frac{\partial p}{\partial \hat{p}}\theta_p + \frac{\partial t}{\partial \hat{p}}\theta_t \\
\theta_{\hat{t}} &= \frac{\partial x}{\partial \hat{t}}\theta_x + \frac{\partial p}{\partial \hat{t}}\theta_p + \frac{\partial t}{\partial \hat{t}}\theta_t
\end{aligned}
\label{generalTransf}
\end{equation}
If we are changing space-time coordinates, then space and time are not functions of momentum. With that in mind, we have:
\begin{equation}
\begin{aligned}
\theta_{\hat{x}} &= \frac{\partial x}{\partial \hat{x}}p + \frac{\partial p}{\partial \hat{x}}0 - \frac{\partial t}{\partial \hat{x}}H = \hat{p} \\
\theta_{\hat{p}} &= 0 p + \frac{\partial p}{\partial \hat{p}}0 - 0 H = 0 \\
\theta_{\hat{t}} &= \frac{\partial x}{\partial \hat{t}}p + \frac{\partial p}{\partial \hat{t}}0 - \frac{\partial t}{\partial \hat{t}}H = -\hat{H}
\end{aligned}
\end{equation}
Which gives us:
\begin{equation}
\begin{aligned}
\hat{p} &= \frac{\partial x}{\partial \hat{x}}p - \frac{\partial t}{\partial \hat{x}}H \\
-\hat{H} &= \frac{\partial x}{\partial \hat{t}}p  - \frac{\partial t}{\partial \hat{t}}H
\end{aligned}
\label{HamiltonianTransf}
\end{equation}

Technically, there are a couple of problems we swept under the rug. We are still using time as the evolution parameter so not every coordinate transformation is possible. For example, imagine switching time with space:
\begin{equation}
\begin{aligned}
\hat{x} &= t \\
\hat{t} &= x
\end{aligned}
\end{equation}
or creating a polar coordinate between space and time:
\begin{equation}
\begin{aligned}
\hat{x} &= \tan^{-1}(\dfrac{x}{t}) \\
\hat{t} &= \sqrt{x^2 + t^2}
\end{aligned}
\end{equation}
Apart the fact that they do not make physical sense, a total derivative in time along all trajectories is not possible. In these cases Hamilton's equation \eqref{Hamilton} do make sense because $\hat{t}$ can no longer be used as the parameter of the evolution.

If the change in space-time coordinate does make sense, we can use \eqref{HamiltonianTransf} to change momentum and Hamiltonian to the new coordinates. Let's try doing that for Galilean transformations.

\section{Galilean transformations revisited}
If we take the transformations rules for a Galilean boost use them into \eqref{HamiltonianTransf} we have:
\begin{equation}
\begin{aligned}
\hat{x} &= x + v_0 t \\
\hat{t} &= t \\
\hat{p} &= p \\
\hat{H} &= v_0 p + H
\end{aligned}
\label{galTransf}
\end{equation}
since $\frac{\partial x}{\partial \hat{x}} = \frac{\partial (\hat{x} - v_0 t)}{\partial \hat{x}} = 1$ and $\frac{\partial x}{\partial \hat{t}} = \frac{\partial (\hat{x} - v_0 t)}{\partial \hat{t}} = - v_0$. Note that these equations are valid for any Hamiltonian.

Let's see what happens if we use the Hamiltonian for a Galilean free particle $H=\dfrac{p^2}{2m}$:
\begin{equation}
\begin{aligned}
\hat{H} &= v_0 p + \dfrac{p^2}{2m} = \dfrac{p^2}{2m} + \dfrac{2mv_0 p}{2m} + \dfrac{m^2 v_0^2}{2m} - \dfrac{m^2 v_0^2}{2m} \\
&= \dfrac{(p+mv_0)^2}{2m} - \dfrac{1}{2} m v_0^2
\end{aligned}
\label{galTransf}
\end{equation}
We kind of get what we expect since:
\begin{equation}
\begin{aligned}
\hat{v} &= v + v_0 =  \frac{p}{m} + v_0 \\
\hat{H} &= \dfrac{1}{2} m \hat{v}^2 - \dfrac{1}{2} m v_0^2
\end{aligned}
\label{galTransf}
\end{equation}
But note that $\hat{p} = p$: the momentum does not change. We can set $\hat{p} = p + m v_0$ since changing the momentum by a constant does not change the equations, but it's an ad-hoc change. Also note that the Hamiltonian is not exactly form invariant under the transformation: since it gets a negative constant it is not exactly the same function of position and momentum. The Hamiltonian for a Galilean free particle, then, is not the same for all inertial observers.

\section{Conclusion}

The first important point is that the transformation rules \eqref{HamiltonianTransf} are the ones one encounters in special relativity for the four-momentum. Note that we haven't said whether the speed of light should be conserved therefore those relationships are more fundamental.

The second important point is that the relationship $p=mv$ is not valid in general. It is only valid for inertial frames and after we ``retouch" the transformation rules. We will see a slightly better version in a future post.

The third important point is that we have seen how Galilean mechanics fails to be relativistic. While the Hamiltonian will never be invariant, what we want it to be form-invariant. That is: for all inertial observers, the expression of the Hamiltonian for the free particle in terms of position, momentum and time must be the same. The problems is that, physically, all inertial systems are the same to us. If the Hamiltonian has a different form for the different inertial systems, then they are not indistinguishable.

This is exactly where a lot of people get confused: the Hamiltonian is not invariant (i.e. the energy is different for different observers, as position, time or momentum) yet it can be form-invariant (i.e. its expression in terms of position, time and momentum is the same). And maybe we should look into it in the next post.

\end{document}

### 2017-09-14 Blog/11-HamiltoniansAndCoordinateChanges.tex
Final review of next blog post

@@
 
-\section{Hamiltonians and coordinate changes}
+\section{How Hamiltonians vary under coordinate changes}
 
@@
 
-In the previous post we saw how momentum changes as covariant components and that keeps the Hamiltonian equations unchanged under coordinate transformations.
-We have also seen, though, that under coordinate transformation that mix time and space we have a problem.
+In the previous post we saw how momentum varies as covariant components and that keeps the Hamiltonian equations unchanged under coordinate transformations.
+We have also seen, though, that under coordinate transformations that mix time and space we have a problem.
 
 What we'll show in this post is that if we impose that Hamilton's equations remain unchanged for all coordinate transformations then the Hamiltonian itself is not invariant but it changes as the covariant time component of a covector.
-That is, in the same way we group space and time $(x, t)$, we also group momentum and Hamiltonian into a covector of components $(p, -H)$.
+That is, in the same way we group space and time $(x, t)$, we also group momentum and the Hamiltonian into a covector of components $(p, -H)$.
 
-Let's see how this work.
+Let's see how this works.
 
@@
 At least, not one that generalizes well.
-Let's at least intuitively try to understand why the Hamiltonian cannot possibly be invariant.
+Let's first intuitively try to understand why the Hamiltonian cannot possibly be invariant.
 
@@
 Note, though, that they always appear inverted: when one is in the numerator, the other is in the denominator and vice-versa.
-That is way they have to change in the opposite way: if space is contravariant while momentum is covariant then the equations stay the same.
+That is why they have to change in the opposite way: if space is contravariant while momentum is covariant then the equations stay the same.
 
 If we change time (e.g.
-imagine changing units by multiplying by a constant) and nothing else changes than the equations change as well.
+imagine changing units by multiplying by a constant) and nothing else changes then the equations change as well (e.g.
+the constant appears only on one side).
 Note, though, that the Hamiltonian appears inverted on the opposite side of the equation: time is always in the denominator while the Hamiltonian is always in the numerator.
@@
 
-In a previous post, we also saw how Hamilton's equations for a single degree of freedom can also be written as:
+To find the transformation law, in a previous post we saw how Hamilton's equations for a single degree of freedom can also be written as:
 \begin{equation}
@@
 \end{equation}
-If we are changing space-time coordinates, then space and time are not functions of momentum. With that in mind, we have:
+If we are changing space-time coordinates $\hat{x}(x,t)$ and $\hat{t}(x,t)$, then space and time are not functions of momentum. With that in mind, we have:
 \begin{equation}
@@
 \end{equation}
-Apart the fact that they do not make physical sense, a total derivative in time along all trajectories is not possible.
-In these cases Hamilton's equation \eqref{Hamilton} do make sense because $\hat{t}$ can no longer be used as the parameter of the evolution.
+Apart from the fact that they do not make physical sense, a total derivative in time along all trajectories is not possible in the new coordinates.
+In these cases Hamilton's equations \eqref{Hamilton} do not make sense because $\hat{t}$ can no longer be used as the parameter of the evolution.
 
-If the change in space-time coordinate does make sense, we can use \eqref{HamiltonianTransf} to change momentum and Hamiltonian to the new coordinates.
+If the change in space-time coordinates does make sense, we can use \eqref{HamiltonianTransf} to change momentum and the Hamiltonian to the new coordinates.
 Let's try doing that for Galilean transformations.
@@
 \section{Galilean transformations revisited}
-If we take the transformations rules for a Galilean boost use them into \eqref{HamiltonianTransf} we have:
+If we take the transformation rules for a Galilean boost and plug them into \eqref{HamiltonianTransf} we have:
 \begin{equation}
@@
 
-The first important point is that the transformation rules \eqref{HamiltonianTransf} are the ones one encounters in special relativity for the four-momentum.
-Note that we haven't said whether the speed of light should be conserved therefore those relationships are more fundamental.
+The first important point is that the transformation rules \eqref{HamiltonianTransf} are essentially the ones one encounters in special relativity for the four-momentum.
+Note that we haven't said whether the speed of light should be invariant therefore these transformation rules are more fundamental.
 
@@
 It is only valid for inertial frames and after we ``retouch" the transformation rules.
-We will see a slightly better version in a future post.
+If you studied special relativity, note that the similar relationships between four-momentum and four-velocity, instead, hold more generally.
 
-The third important point is that we have seen how Galilean mechanics fails to be relativistic.
-While the Hamiltonian will never be invariant, what we want it to be form-invariant.
+The third important point is that we have seen in what way Galilean mechanics fails to be relativistic.
+While the Hamiltonian will never be invariant, we want it to be form-invariant.
 That is: for all inertial observers, the expression of the Hamiltonian for the free particle in terms of position, momentum and time must be the same.
-The problems is that, physically, all inertial systems are the same to us.
+Physically all inertial systems are the same to us.
 If the Hamiltonian has a different form for the different inertial systems, then they are not indistinguishable.
@@
 This is exactly where a lot of people get confused: the Hamiltonian is not invariant (i.e.
-the energy is different for different observers, as position, time or momentum) yet it can be form-invariant (i.e.
+the energy is different for different observers, as are position, time and momentum) yet it can be form-invariant (i.e.
 its expression in terms of position, time and momentum is the same).
-And maybe we should look into it in the next post.
 

### 2017-09-21 Blog/12-ExtendedHamiltonians.tex
Draft on new blog post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Hamiltonian mechanics on the extended phase space}

TL;DR The most complete version of Hamiltonian mechanics lives on the extended phase space that includes position, momentum, time and energy.

In this post we look at the most complete version of Hamiltonian mechanics. It is kind of a shame that this is not the version taught because it is the one that works best mathematically and is the one that allows to make parallels to quantum mechanics and special relativity. The general idea is that, instead of just having phase space made of position and momentum, we extend it to include time and energy. This was we can separate time as a coordinate (so that we can do all the coordinate changes we want) and time as the evolution parameter (so that we can keep describing the evolution no matter the coordinates).

Let's see how this works.

\section{Hamiltonian mechanics in the extended phase space}

We consider the space formed by $(x^i,t,p_j,E)$. Each point particle will have trajectories $(x^i(s),t(s),p_j(s),E(s))$ where $s$ is the evolution parameter. The equations of motion will be given by:
\begin{equation}
\begin{aligned}
\frac{dx^i}{ds} &= \frac{\partial \mathcal{H}}{\partial p_i} \\
\frac{dt}{ds} &= - \frac{\partial \mathcal{H}}{\partial E} \\
\frac{dp_j}{ds} &= - \frac{\partial \mathcal{H}}{\partial x^j} \\
\frac{dE}{ds} &= \frac{\partial \mathcal{H}}{\partial t} \\
\end{aligned}
\label{ExtendedHamilton}
\end{equation}
where $\mathcal{H}(x^i,t,p_j,E)$ is the extended Hamiltonian. Time and energy have equations that are similar to the ones for position and momentum, except they have a sign difference. Under coordinate transformations $(p_j, -E)$ will change as a covector. This is very easy to see now because we can rewrite the equations as:
\begin{equation}
\begin{aligned}
\frac{dx^i}{ds} &= \frac{\partial \mathcal{H}}{\partial p_i} \\
\frac{dt}{ds} &= \frac{\partial \mathcal{H}}{\partial (-E)} \\
\frac{dp_j}{ds} &= - \frac{\partial \mathcal{H}}{\partial x^j} \\
\frac{d(-E)}{ds} &= - \frac{\partial \mathcal{H}}{\partial t} \\
\end{aligned}
\label{ExtendedHamiltonRewritten}
\end{equation}
So $-E$ behaves exactly like a component of momentum. The evolution parameter $s$ does not change and therefore the extended Hamiltonian does not change either: the extended Hamiltonian is invariant.

So, what is the connection between the standard Hamiltonian and the extended Hamiltonian? Between the standard set of equation and the extended set of equations?

Suppose we have a particular system where we can set $\mathcal{H} = H(x^i,p_j) - E$. We have:
\begin{equation}
\begin{aligned}
\frac{dx^i}{ds} &= \frac{\partial H}{\partial p_i} \\
 \frac{dt}{ds} &= 1 \\
\frac{dp_j}{ds} &= - \frac{\partial H}{\partial x^j} \\
 \frac{dE}{ds} &= 0 \\
\end{aligned}
\label{HamiltonRecovered}
\end{equation}
So $t=s$ and we recover the standard equations. That is: in the special case where we use time as the evolution parameter and the energy does not change in time we recover the standard Hamiltonian mechanics.

Not only the extended Hamiltonian is invariant, but it also is constant throughout the evolution and it is customary to simply set it to zero. In the previous case we have $\mathcal{H} = 0 = H(x^i,p_j) - E$: $E = H(x^i,p_j)$.

We can see that the extended Hamiltonian has two functions: the first is as a potential for the equations for motion; the second as a constrain that links the energy to position, momentum and time.

\section{Conclusion}

This formulation of Hamiltonian mechanics is the most general and, personally, I think it's the correct one as it allows to tie in even more concepts from quantum mechanics and special relativity.

We have seen how Hamiltonian mechanics already has an uncertainty principle, we clearly see here that energy and momentum form a covector and we will see that in extended phase space we can find classical analogues of antiparticles and other ideas that are connected to quantum field theories.

\end{document}

### 2017-10-10 Blog/12-ExtendedHamiltonians.tex
Minor corrections

@@
 In this post we look at the most complete version of Hamiltonian mechanics.
-It is kind of a shame that this is not the version taught because it is the one that works best mathematically and is the one that allows to make parallels to quantum mechanics and special relativity.
-The general idea is that, instead of just having phase space made of position and momentum, we extend it to include time and energy.
-This was we can separate time as a coordinate (so that we can do all the coordinate changes we want) and time as the evolution parameter (so that we can keep describing the evolution no matter the coordinates).
+It is kind of a shame that this is not the version taught because it is the one that works best mathematically and is the one that allows parallels to quantum mechanics and special relativity.
+The general idea is that, instead of just having phase space consisting of position and momentum, we extend it to include time and energy.
+This way we can separate time as a coordinate (so that we can do all the coordinate changes we want) and time as the evolution parameter (so that we can keep describing the evolution no matter the coordinates).
 
@@
 
-\section{Hamiltonian mechanics in the extended phase space}
+\section{Hamilton's equations on the extended phase space}
 
@@
 
-So, what is the connection between the standard Hamiltonian and the extended Hamiltonian? Between the standard set of equation and the extended set of equations?
+So, what is the connection between the standard Hamiltonian and the extended Hamiltonian? Between the standard set of equations and the extended set of equations?
 
@@
 So $t=s$ and we recover the standard equations.
-That is: in the special case where we use time as the evolution parameter and the energy does not change in time we recover the standard Hamiltonian mechanics.
+That is: in the special case where we use time as the evolution parameter and the energy does not change in time we recover standard Hamiltonian mechanics.
 
-Not only the extended Hamiltonian is invariant, but it also is constant throughout the evolution and it is customary to simply set it to zero.
+Not only is the extended Hamiltonian invariant, but it is also constant throughout the evolution and it is customary to simply set it to zero.
 In the previous case we have $\mathcal{H} = 0 = H(x^i,p_j) - E$: $E = H(x^i,p_j)$.
 
-We can see that the extended Hamiltonian has two functions: the first is as a potential for the equations for motion; the second as a constrain that links the energy to position, momentum and time.
+We can see that the extended Hamiltonian has two functions: the first is as a potential for the equations of motion; the second as a constraint that links the energy to position, momentum and time.
 
@@
 
-This formulation of Hamiltonian mechanics is the most general and, personally, I think it's the correct one as it allows to tie in even more concepts from quantum mechanics and special relativity.
+This formulation of Hamiltonian mechanics is the most general and, personally, I think is ultimately the correct one as it allows more concepts from quantum mechanics and special relativity to be tied in.
 

### 2017-10-10 Blog/13-FreeParticleExtendedHamiltonian.tex
New blog post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Classical anti-particles}

TL;DR Anti-particle states are those for which the parametrization of the trajectory in phase-space is anti-aligned with time (i.e. it decreases as time increases).

In a previous post we introduced the extended Hamilton's equations. In the next few posts we will see that they allow to create parallels not only to special relativity, but also quantum field theory. That is: many quantum concepts have classical counterparts.

The first thing I want to show you is that there is such a thing as classical anti-particle. One of the key feature of the extended Hamilton equations is that we have separated the time coordinate $t$ from the evolution parameter $s$. The idea is that $t$ and $s$ can be anti-aligned, that is if $t$ decreases when $s$ increases. Those exactly correspond to anti-particles.

Let's look at the details.

\section{Time coordinate and evolution parameter}

The spatial trajectory of a particle in terms of the evolutions parameter $s$ can be written as $(x^i(s), t(s))$. But if $t$ is a valid time coordinate, we must have only one position for each moment in time. That is: the trajectory cannot pass twice at a given $t$. For each value of $t$ there is one only one $s$. The function $t(s)$ must be invertible.

But if $t(s)$ is invertible, it must be strictly monotonic, either decreasing or increasing, along the whole trajectory. Now consider the following part of the extended Hamiltonian equations:
\begin{equation}
\begin{aligned}
\frac{dt}{ds} &= - \frac{\partial \mathcal{H}(x^i,t,p_j,E)}{\partial E} \\
\end{aligned}
\label{ExtendedHamiltonTime}
\end{equation}
The extended Hamiltonian $\mathcal{H}$ will determine whether $t(s)$ is increasing or decreasing. That is, for each state in the extended phase space $(x^i,t,p_j,E)$ we will know whether $t(s)$ increases or decreases in that neighborhood.

Let's call ``standard particle states" those for which $t(s)$ increases and ``antiparticle states" those for which $t(s)$ decreases. Note that a trajectory cannot cross those two groups: if it did $t(s)$ would not be strictly monotonic. So standard particle states can only evolve into other standard particle states and anti-particle states can only evolve into anti-particle states.

What we call particle or anti-particles is arbitrary: if we invert the sign of $\mathcal{H}$, then we invert the direction of $s$ with respect to the other variables. But the fact that there are these two groups is something that is invariant under all parametrization.

Now, this may all be nice mathematically, but physically we think of anti-particle as having opposite charge. How does that come in? The general idea is that the conservative forces, like electromagnetism, are such that we can write:
\begin{equation}
\begin{aligned}
m\frac{d^2x^\alpha}{ds^2} &= F^\alpha_\beta \frac{dx^\beta}{ds} \\
\end{aligned}
\label{NewtonHamilton}
\end{equation}
That is: they are linear functions of the four-velocity. Now suppose we re-express those in terms of a parametrization $\tau$ that always increases with time. For standard particles we have $s = \tau$:
\begin{equation}
\begin{aligned}
	m\frac{d^2x^\alpha}{d\tau^2} &= F^\alpha_\beta \frac{dx^\beta}{d\tau} \\
\end{aligned}
\label{NewtonHamiltonParticles}
\end{equation}
For anti-particles we have $s = - \tau$:
\begin{equation}
\begin{aligned}
m\frac{d^2x^\alpha}{d(-\tau)^2} &= F^\alpha_\beta \frac{dx^\beta}{d(-\tau)} \\
m\frac{d^2x^\alpha}{d(\tau)^2} &= - F^\alpha_\beta \frac{dx^\beta}{d\tau} \\
\end{aligned}
\label{NewtonHamiltonAntiParticles}
\end{equation}
That is: the second derivative does not change sign but the first derivative does. Effectively, the force is opposite for the anti-particles.

\section{Conclusion}

If you have studied quantum field theory and you have seen Feynman diagrams, you see that anti-particles are drawn like particles with arrows in the opposite direction. The direction is not of time: it's not that they are traveling backward in time. It's the direction of the parametrization: if we use the Hamilton equations to evolve the state of the anti-particle, we reconstruct the past instead of predicting the future.

Anti-particles are not an exclusive feature of quantum field theory. They are a feature of Hamiltonian mechanics. The problem is that you need the extended phase-space to be able to see them, and, since this is not what is typically done, you don't. On the other hand, it is impossible to write a relativistic quantum field theory without working on the extended phase space, so you cannot hide them there.

\end{document}

### 2017-10-23 Blog/13-ClassicalAntiParticles.tex
New blog post

--- 2017-10-10 Blog/13-FreeParticleExtendedHamiltonian.tex
@@
 
-\section{Classical anti-particles}
+\section{Classical antiparticles}
 
-TL;DR Anti-particle states are those for which the parametrization of the trajectory in phase-space is anti-aligned with time (i.e. it decreases as time increases).
+TL;DR Antiparticle states are those for which time is a decreasing function $t(s)$ of the trajectory parametrization instead of an increasing function (i.e.
+time and trajectory parametrization are anti-aligned).
 
-In a previous post we introduced the extended Hamilton's equations.
-In the next few posts we will see that they allow to create parallels not only to special relativity, but also quantum field theory.
+In a previous post we introduced the extended Hamiltonian equations.
+In the next few posts we will see that they allow us to create parallels not only to special relativity, but also quantum field theory.
 That is: many quantum concepts have classical counterparts.
 
-The first thing I want to show you is that there is such a thing as classical anti-particle.
-One of the key feature of the extended Hamilton equations is that we have separated the time coordinate $t$ from the evolution parameter $s$.
+The first thing I want to show you is that there is such a thing as classical antiparticles.
+One of the key features of the extended Hamiltonian equations is that we have separated the time coordinate $t$ from the evolution parameter $s$.
 The idea is that $t$ and $s$ can be anti-aligned, that is if $t$ decreases when $s$ increases.
-Those exactly correspond to anti-particles.
+Those states correspond to anti-particles.
 
@@
 
-The spatial trajectory of a particle in terms of the evolutions parameter $s$ can be written as $(x^i(s), t(s))$.
+The trajectory of a particle in terms of the evolution parameter $s$ can be written as $(x^i(s), t(s))$.
 But if $t$ is a valid time coordinate, we must have only one position for each moment in time.
@@
 Note that a trajectory cannot cross those two groups: if it did $t(s)$ would not be strictly monotonic.
-So standard particle states can only evolve into other standard particle states and anti-particle states can only evolve into anti-particle states.
+So standard particle states can only evolve into other standard particle states and antiparticle states can only evolve into other antiparticle states.
 
-What we call particle or anti-particles is arbitrary: if we invert the sign of $\mathcal{H}$, then we invert the direction of $s$ with respect to the other variables.
-But the fact that there are these two groups is something that is invariant under all parametrization.
+What we call particles or antiparticles is arbitrary as the parametrization is not physical.
+In fact, if we invert the sign of $\mathcal{H}$, then we invert the direction of $s$ with respect to the other variables.
+But the fact that there are these two groups is something that is invariant under all parametrizations.
 
-Now, this may all be nice mathematically, but physically we think of anti-particle as having opposite charge.
+Now, this may all be nice mathematically, but physically we think of antiparticles as having opposite charge.
 How does that come in?
@@
 \end{equation}
-That is: they are linear functions of the four-velocity.
+That is: the four-acceleration is a linear function of the four-velocity.
 Now suppose we re-express those in terms of a parametrization $\tau$ that always increases with time.
@@
 \end{equation}
-For anti-particles we have $s = - \tau$:
+For antiparticles we have $s = - \tau$:
 \begin{equation}
 \begin{aligned}
-m\frac{d^2x^\alpha}{d(-\tau)^2} &= F^\alpha_\beta \frac{dx^\beta}{d(-\tau)} \\
-m\frac{d^2x^\alpha}{d(\tau)^2} &= - F^\alpha_\beta \frac{dx^\beta}{d\tau} \\
+m\frac{d^2x^\alpha}{(d(-\tau))^2} &= F^\alpha_\beta \frac{dx^\beta}{d(-\tau)} \\
+m\frac{d^2x^\alpha}{d\tau^2} &= - F^\alpha_\beta \frac{dx^\beta}{d\tau} \\
 \end{aligned}
@@
 \end{equation}
-That is: the second derivative does not change sign but the first derivative does. Effectively, the force is opposite for the anti-particles.
+That is: the second derivative does not change sign but the first derivative does. Effectively, the force is opposite for the antiparticles.
 
@@
 
-If you have studied quantum field theory and you have seen Feynman diagrams, you see that anti-particles are drawn like particles with arrows in the opposite direction.
-The direction is not of time: it's not that they are traveling backward in time.
-It's the direction of the parametrization: if we use the Hamilton equations to evolve the state of the anti-particle, we reconstruct the past instead of predicting the future.
+If you have studied quantum field theory and you have seen Feynman diagrams, you see that antiparticles are drawn like particles with arrows in the opposite direction.
+That direction is not the direction of time: it's not that antiparticles are traveling backward in time.
+It's the direction of the parametrization: if we use the extended Hamiltonian equations to evolve the state of the antiparticle forward in the parametrization, we reconstruct the past instead of predicting the future.
 
-Anti-particles are not an exclusive feature of quantum field theory.
+Antiparticles are not an exclusive feature of quantum field theory.
 They are a feature of Hamiltonian mechanics.
-The problem is that you need the extended phase-space to be able to see them, and, since this is not what is typically done, you don't.
+The problem is that you need the extended phase space to be able to see them, and, since this is not what is typically done in classical mechanics, you don't.
 On the other hand, it is impossible to write a relativistic quantum field theory without working on the extended phase space, so you cannot hide them there.

### 2017-10-23 Blog/14-FreeParticleExtendedHamiltonian.tex
New blog post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Classical Klein-Gordon}

TL;DR The extended Hamiltonian for a classical free particle is the classical version of the Klein-Gordon equation.

After having seen the classical version of antiparticle we will see the classical version of the Klein-Gordon equation. This is the essentially the extended Hamiltonian for the free particle.

Let's look at the details.

\section{Extended Hamiltonian for the free particle}

The extended Hamiltonian for a free particle is:
\begin{equation}
\begin{aligned}
\mathcal{H}(x^i,t,p_j,E) = \frac{1}{2m} \left(p^2 - \left(\frac{E^2}{c^2}\right)\right)  + \frac{1}{2} mc^2\\
\end{aligned}
\label{ExtendedHamiltonFree}
\end{equation}

Remember that, for valid physical states, we have $\mathcal{H}=0$. Therefore:
\begin{equation}
\begin{aligned}
\left(p^2 - \left(\frac{E^2}{c^2}\right)\right) = - m^2c^2\\
\end{aligned}
\label{FourMomentum}
\end{equation}
If you studied special relativity, you recognized the norm of the four-momentum. Moreover:
\begin{equation}
\begin{aligned}
\frac{E^2}{c^2} &= p^2 + m^2c^2\\
E &= \pm c \sqrt{p^2 + m^2c^2} \\
\end{aligned}
\label{HamiltonFree}
\end{equation}
This is the energy as a function of momentum: it the relativistic Hamiltonian for a free particle. But in this case the energy can both be positive or negative. Note that:
\begin{equation}
\begin{aligned}
\frac{dt}{ds} = - \frac{\partial \mathcal{H}}{\partial E} = \frac{E}{mc^2}
\end{aligned}
\label{TimeEvolution}
\end{equation}
So if the energy is negative, then the parametrization is anti-aligned to the time: $t$ decreases as $s$ increases. As we saw in the previous post, these correspond to anti-particle states.

So, not only the extended Hamiltonian recovers relativistic free-particles, it also recovers anti-particles and it is has much nicer form than the relativistic Hamiltonian (i.e. it is quadratic and does not have any square root).

\section{Relation to Klein-Gordon}

If you use the standard classical to quantum substitutions $\left( p_i, -E \right)$ to $\left(-\imath \hbar \frac{\partial}{\partial x^i}, -\imath \hbar \frac{\partial}{\partial t} \right)$ we have:
\begin{equation}
\begin{aligned}
\mathcal{H} = &\frac{1}{2m} \left(- \hbar^2 \nabla^2 + \hbar^2 \left(\frac{\partial^2}{c^2\partial t^2}\right)\right)  + \frac{1}{2} mc^2 = 0\\
&\frac{1}{c^2}\frac{\partial^2}{\partial t^2}- \nabla^2 + \frac{m^2 c^2}{\hbar^2}  = 0\\
\end{aligned}
\label{KleinGordon}
\end{equation}
If you studied quantum field theory, that should look familiar: if we add a $\psi$ it's the Klein-Gordon equation. (Note: the mass term is positive as we are using the relativity metric convention $(-,+,+,+)$ instead of the particle physics metric convention $(+, -, -, -)$).

So why should we add $\psi$? There is good reason to do this, even in the classical setting.

Suppose you have a phase space distribution $(x^i,t,p_j,E)$. This would correspond to a distribution of matter over space, time, momentum and energy. Consider $\mathcal{H} \rho$, the product between the extended Hamiltonian and the distribution. As we saw, $\mathcal{H}=0$ is a constraint between energy and the other variables so that only states that satisfy that equation are physical. Therefore $\rho$ can only be different than zero for those states where $\mathcal{H}$ is zero. But this means that $\mathcal{H} \rho = 0$ everywhere. In other words: $\mathcal{H} \rho = 0$ can be used as a constraint to find the valid distribution.

Now, in the quantum case states are always ensembles: they are always distributions. Therefore we use the constraint $\mathcal{H}\psi = 0$. Which, in the case of the free particle, corresponds to the Klein-Gordon equation.

\section{Conclusion}

The extended Hamiltonian for a free particle has a much nicer form than the relativistic Hamiltonian, and it ties in more with quantum concepts. The Klein-Gordon equation is the extended Hamiltonian for a free particle acting on the distribution. The only thing it does is constrain the state so that the norm of the four-momentum is $-m^2c^2$.

\end{document}

### 2017-11-14 Blog/14-FreeParticleExtendedHamiltonian.tex
Minor corrections

@@
 
-After having seen the classical version of antiparticle we will see the classical version of the Klein-Gordon equation. This is the essentially the extended Hamiltonian for the free particle.
+After having seen the classical version of antiparticles we will see the classical version of the Klein-Gordon equation. This is the essentially the extended Hamiltonian for the free particle.
 
@@
 
-Suppose you have a phase space distribution $(x^i,t,p_j,E)$.
+Suppose you have a phase space distribution $\rho(x^i,t,p_j,E)$.
 This would correspond to a distribution of matter over space, time, momentum and energy.
@@
 But this means that $\mathcal{H} \rho = 0$ everywhere.
-In other words: $\mathcal{H} \rho = 0$ can be used as a constraint to find the valid distribution.
+In other words: $\mathcal{H} \rho = 0$ can be used as a constraint to find physically meaningful distributions over the extended phase space.
 
@@
 
-The extended Hamiltonian for a free particle has a much nicer form than the relativistic Hamiltonian, and it ties in more with quantum concepts.
+The extended Hamiltonian for a free particle has a much nicer form than the relativistic Hamiltonian one finds in standard textbooks, and it ties in more with quantum concepts.
 The Klein-Gordon equation is the extended Hamiltonian for a free particle acting on the distribution.

### 2017-11-14 Blog/15-ClassicalSpin.tex
New post

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Classical spin}

TL;DR The Hamiltonian description of a direction in space is the classical version of the spin of a particle.

Now we turn our attention to the classical version of another quantum concept: spin. What we show here is that Hamiltonian motion of a spatial direction is qualitatively the same as the evolution of spin 1/2.

Let's look at the details.

\section{Phase space for a direction in space}

Suppose we have a vector $\vec{S} = \{S_x, S_y, S_z\}$ and suppose we want to give a Hamiltonian description to its direction. The phase space we want to consider, then, is the space of all possible directions, which we can think of the sphere of unitary radius. The norm of the vector $|\vec{S}|$, instead, is a parameter for the system, much in the same way that mass is parameter for a massive particle. So, for convenience, we will assume that $\vec{S}$ is unitary.

The first order of business is to find a set of conjugate variables. Recall that $x$ and $p_x$ are conjugate because the area $dx \wedge dp_x$ does not change under coordinate (i.e. unit) transformations. Note that the solid angle $d\Omega$ is already a quantity that is independent of coordinates so we can write:
\begin{equation}
\begin{aligned}
d\Omega &= \sin \psi d\psi d\theta \\
&= - d(\cos \psi) d\theta \\
&= - d(\cos \psi) \wedge d\theta \\
&=  d\theta \wedge d(\cos \psi)
\end{aligned}
\label{DirectionDOF}
\end{equation}
The minus sign is removed since the wedge product is anti-symmetric (i.e. $a \wedge b = - b \wedge a$). Note that if we take $\cos \psi$ is the component of the vector along the vertical direction. So our conjugate variables are $\theta^{xy}$, the azimuthal angle, and $S_z$, the vertical component.
%picture https://upload.wikimedia.org/wikipedia/commons/d/dc/3D_Spherical_2.svg

In terms of the conjugate variable, we have:
\begin{equation}
\begin{aligned}
S_x &= \sqrt{1-S_z^2} \cos \theta^{xy} \\
S_y &= \sqrt{1-S_z^2} \sin \theta^{xy} \\
S_z &= S_z\\
\end{aligned}
\label{SpinComponents}
\end{equation}

We can calculate the Poisson brackets:
\begin{equation}
\begin{aligned}
\{S_x, S_y\} &= \frac{\partial S_x}{\partial \theta^{xy}}\frac{\partial S_y}{\partial S_z} - \frac{\partial S_y}{\partial \theta^{xy}}\frac{\partial S_x}{\partial S_z} \\
&= -\sin \theta^{xy} \sqrt{1-S_z^2} \sin \theta^{xy} (- \frac{S_z}{\sqrt{1-S_z^2}})  \\ 
&\; - \cos \theta^{xy} \sqrt{1-S_z^2} \cos \theta^{xy} (- \frac{S_z}{\sqrt{1-S_z^2}}) \\
&= \sin^2 \theta^{xy} S_z +\cos^2 \theta^{xy} S_z \\
&= S_z \\
\{S_y, S_z\} &= \frac{\partial S_y}{\partial \theta^{xy}}\frac{\partial S_z}{\partial S_z} - \frac{\partial S_z}{\partial \theta^{xy}}\frac{\partial S_y}{\partial S_z} \\
&= \cos \theta^{xy} \sqrt{1-S_z^2} \cdot 1 - 0 \\
&= S_x \\
\{S_z, S_x\} &= \frac{\partial S_z}{\partial \theta^{xy}}\frac{\partial S_x}{\partial S_z} - \frac{\partial S_x}{\partial \theta^{xy}}\frac{\partial S_z}{\partial S_z} \\
&= 0 - (-\sin \theta^{xy} \sqrt{1-S_z^2}) \cdot 1 \\
&= S_y \\
\end{aligned}
\label{SpinBrackets}
\end{equation}
Note that these are analogous to the same relationships one has with spin components and commutators in quantum mechanics.

\section{Relation to spin and magnetic field}

Suppose we have the following Hamiltonian $H=\mu\vec{B}\cdot\vec{S}=\mu B^i S_i$, which is formally identical to the Hamiltonian for a spin 1/2 system in a uniform magnetic field. We can calculate the evolution of the vector components using the standard Hamiltonian techniques. For simplicity, let's assume our ``field" $B$ is oriented along the $z$ axis, that is $H= \mu B^z S_z$. We have:
\begin{equation}
\begin{aligned}
\frac{dS_x}{dt} &= \{S_x, H\} = \mu B^z \{S_x, S_z\} = - \mu B^z S_y \\
\frac{dS_y}{dt} &= \{S_y, H\} = \mu B^z \{S_y, S_z\} = \mu B^z S_x \\
\frac{dS_z}{dt} &= \{S_z, H\} = \mu B^z \{S_z, S_z\} = 0 \\
\end{aligned}
\label{SpinEquations}
\end{equation}

Integrating, we have:
\begin{equation}
\begin{aligned}
S_x &= S^0_x \cos \omega t - S^0_y \sin \omega t \\
S_y &= S^0_x \sin \omega t + S^0_y \cos \omega t \\
S_z &= S^0_z \\
\omega &= \mu B^z
\end{aligned}
\label{SpinPrecession}
\end{equation}

The vector $\vec{S}$ precesses about the axis determined by $\vec{B}$, and in the plane orthogonal to it, with angular velocity $\mu B^z$. This is exactly what spin does in a magnetic field.

\section{Conclusion}

As we have seen, we can construct a classical analogue of spin just by looking at the Hamiltonian evolution of spatial directions. Like in the case of position and momentum, we have two conjugate variables and we have the same set of relationship and equations between them.

Another thing to note. In Hamiltonian mechanics the conjugate variables are a pair. If the spatial direction is an independent degree of freedom then it must be described by a pair of variable. Note that a direction in an $n$ dimensional space is always described by a set of $n-1$ variables. Therefore the only direction we can describe in Hamiltonian mechanics as an independent degrees of freedom is a direction in a three dimensional space.

\end{document}

### 2017-11-15 Blog/15-ClassicalSpin.tex
Minor corrections

@@
 Suppose we have a vector $\vec{S} = \{S_x, S_y, S_z\}$ and suppose we want to give a Hamiltonian description to its direction.
-The phase space we want to consider, then, is the space of all possible directions, which we can think of the sphere of unitary radius.
-The norm of the vector $|\vec{S}|$, instead, is a parameter for the system, much in the same way that mass is parameter for a massive particle.
+The phase space we want to consider, then, is the space of all possible directions, which we can think of as the sphere of unitary radius.
+The norm of the vector $|\vec{S}|$, instead, is a parameter for the system, much in the same way that mass is a parameter for a massive particle.
 So, for convenience, we will assume that $\vec{S}$ is unitary.
@@
 $a \wedge b = - b \wedge a$).
-Note that if we take $\cos \psi$ is the component of the vector along the vertical direction.
-So our conjugate variables are $\theta^{xy}$, the azimuthal angle, and $S_z$, the vertical component.
+Note that  $\cos \psi$ is the component of the vector along the vertical direction.
+So our conjugate variables are $\theta^{xy}$, the angle on the $(x, y)$ plane, and $S_z$, the vertical component.
 %picture https://upload.wikimedia.org/wikipedia/commons/d/dc/3D_Spherical_2.svg
@@
 \begin{aligned}
-S_x &= \sqrt{1-S_z^2} \cos \theta^{xy} \\
-S_y &= \sqrt{1-S_z^2} \sin \theta^{xy} \\
+S_x &= \cos \theta^{xy} \sqrt{1-S_z^2} \\
+S_y &= \sin \theta^{xy} \sqrt{1-S_z^2} \\
 S_z &= S_z\\
@@
 We can calculate the evolution of the vector components using the standard Hamiltonian techniques.
-For simplicity, let's assume our ``field" $B$ is oriented along the $z$ axis, that is $H= \mu B^z S_z$.
+For simplicity, let's assume our ``field" $\vec{B}$ is oriented along the $z$ axis, that is $H= \mu B^z S_z$.
 We have:
@@
 As we have seen, we can construct a classical analogue of spin just by looking at the Hamiltonian evolution of spatial directions.
-Like in the case of position and momentum, we have two conjugate variables and we have the same set of relationship and equations between them.
+Like in the case of position and momentum, we have two conjugate variables and we have the same set of relationships and equations between them.
 
@@
 In Hamiltonian mechanics the conjugate variables are a pair.
-If the spatial direction is an independent degree of freedom then it must be described by a pair of variable.
+If the spatial direction is an independent degree of freedom then it must be described by a pair of variables.
 Note that a direction in an $n$ dimensional space is always described by a set of $n-1$ variables.
-Therefore the only direction we can describe in Hamiltonian mechanics as an independent degrees of freedom is a direction in a three dimensional space.
+Therefore the only direction we can describe in Hamiltonian mechanics as an independent degree of freedom is a direction in a three dimensional space since we must have $n-1 = 2$.
 

### 2017-11-25 Articles/TopologyAndExperimentalDistinguishability/article.tex
Adding two remarks

@@
 
+\begin{rem}
+	Any scientific measurement can be formulated as an observation in the form above.
+For example, \emph{``the mass of the electron is 510,999 $\pm$ 0.5 eV"} places the value within the open set $U=(510,998.5 , 510,999.5)$.
+\end{rem}
+
 \begin{lem}
@@
 We need to show that both definitions lead to the same mathematical object.
+
+\begin{rem}
+	Any scientific law can be seen as the relationship between experimentally distinguishable elements.
+For example, $F=ma$ tells us that the product of the measurements for mass and acceleration will be equivalent to a measurement of the force.
+\end{rem}
 

### 2017-11-25 Blog/16-PoissonBrackets.tex
Post on Poisson brackets

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{What are Poisson brackets?}

TL;DR The Poisson bracket tells how a quantity changes under a transformation generated by another. It also tells us the state count of a cell of phase space identified by the two variables.

Another operator of particular importance in Hamiltonian mechanics is the Poisson bracket: $\{f,g\}=\frac{\partial f}{\partial x}\frac{\partial g}{\partial p} - \frac{\partial f}{\partial p} \frac{\partial g}{\partial x}$. Its quantum equivalent is the commutator $\frac{[F, G]}{\imath \hbar} = \frac{FG - GF}{\imath \hbar}$. But what does the Poisson bracket signify geometically and physically?

\section{Poisson bracket and change under a generated transformation}

Recall that given $H(x,p)$, Hamilton's equations are:
\begin{equation}
\begin{aligned}
\frac{dx}{dt} &= \frac{\partial H}{\partial p}  \\
\frac{dp}{dt} &= - \frac{\partial H}{\partial x}
\end{aligned}
\label{Hamilton}
\end{equation}

We can generalize these equations so that we can use them for a generic transformation that preserves areas in phase space, not just time evolution. We use a generic parameter $a$ and generic function $A(x,p)$ and we say that $A$ generates the transformation parametrized by $a$ if we have:
\begin{equation}
\begin{aligned}
\frac{dx}{da} &= \frac{\partial A}{\partial p}  \\
\frac{dp}{da} &= - \frac{\partial A}{\partial x}
\end{aligned}
\label{GeneratedTransformation}
\end{equation}
For example, if we have $A=p$ we have:
\begin{equation}
\begin{aligned}
\frac{dx}{da} &= 1  \\
\frac{dp}{da} &= 0
\end{aligned}
\label{GeneratedTranslation}
\end{equation}
That is, the generated transformation is really the translation and the parameter $a$ is the travelled distance. Note that any quantity that is a function of position and momentum is the generator of some area preserving transformation (i.e. it can be used in Hamilton's equations in place of the Hamiltonian) and any infinitesimal transformation that preserves area in phase space has a generator that is a function of position and momentum.

Now we ask: how does a generic function $f(x,p)$ change during this transformation? We have:
\begin{equation}
\begin{aligned}
\frac{df(x(a),p(a))}{da} &= \frac{\partial f}{\partial x}\frac{dx}{da} + \frac{\partial f}{\partial p}\frac{dp}{da} \\
&=\frac{\partial f}{\partial x} \frac{\partial A}{\partial p} - \frac{\partial f}{\partial p}\frac{\partial A}{\partial x}
&=\{f,A\}
\end{aligned}
\label{PoissonGenerator}
\end{equation}
That is: the derivative of a quantity $f$ with respect to the trajectory parameter $a$ is the Poisson bracket of the quantity $f$ with the generator $A$.

With that in mind, the relationship $\{x, p\} = 1 = \frac{dx}{dx}$ means that $p$ generates the translation along $x$ since the change of $x$ along the transformation is $1$. The equation $\frac{df}{dt} = \{f, H\}$ makes sense because $H$ generates time evolution. Moreover, since $\{A, A\} = 0$, a quantity is always constant during a transformation it generates. For example, energy is conserved in time and momentum does not change under translation.

These relationships are identical in quantum mechanics, except the commutator takes the role of the Poisson braket.

\section{Poisson brackets and areas of phase space}

There is also another more geometrical way to understand the Poisson bracket. This works for classical mechanics and I do not know whether it has an analogue in quantum mechanics.

The idea is the following: suppose you have two functions $f(x,p)$ and $g(x,p)$. Suppose you create the area $df dg$. How does it compare to the area $dx dp$? To answer the question, we simply make a coordinate transformation. The Jacobian will relate the two areas in the following way:
\begin{equation}
\begin{aligned}
df dg &= | J | dx dp = \left| \begin{matrix}
\dfrac{\partial f}{\partial x} & \dfrac{\partial f}{\partial p} \\[2.2ex]
\dfrac{\partial g}{\partial x} & \dfrac{\partial g}{\partial p} \end{matrix} \right| dx dp \\
&= \left( \frac{\partial f}{\partial x} \frac{\partial g}{\partial p} - \frac{\partial f}{\partial p}\frac{\partial g}{\partial x} \right) dx dp = \{f,g\} dx dp
\end{aligned}
\label{PoissonJacobian}
\end{equation}
In other words, the Poisson bracket is the Jacobian of the transformation and it tells us how much the area is bigger or smaller.

Physically, the quantity $dxdp$ tells us how many states we have in that region. The Poisson bracket tells us how many more (or less) states we have in the infinitesimal region $dfdg$.

\section{Conclusion}

We can understand the Poisson bracket $\{f,g\}$ in two geometrical ways. It tells us how the quantity $f$ changes under a transformation generated by $g$, that is a transformation where we are using $g$ in place of the Hamiltonian. It also tells us how many possible states we have in an infinitesimal area $df dg$ in units of $dx dp$.

\end{document}

### 2018-02-13 Blog/17-ClassicalCPT.tex
Adding blog posts

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{CPT-like theorem for classical particles}

TL;DR Inverting the direction of time and space is equivalent to inverting the direction of the parametrization.

In previous posts we have seen the classical equivalent for anti-particles. Here we see something that looks similar to the CPT theorem. Naturally, it's not a complete analogue because we are not really looking at fields at all (just the state of a single particle moving in phase-space), we don't have spin, etc... But the symmetry we get has the same flavor.

As we saw, the classical analogue for anti-particle are simply those states for which the parametrization of the phase-space trajectory is anti-aligned with time. If we flip both space and time, the vectors tangent to the evolution will flip sign and this is equivalent to flipping the sign of the parametrization.

\section{Reflections in space, time and evolution parameter}

Suppose we have a state in the extend phase space $(x^i,t,p_j,E)$. It's change during the evolution will be given by the vector $\frac{d}{ds}(x^i,t,p_j,E) = (\frac{dx^i}{ds},\frac{dt}{ds},\frac{dp_j}{ds},\frac{dE}{ds})$.

Now suppose we perform a spatial reflection. This will send $x^i$ to $-x^i$, multiplying the coordinate by the constant $-1$. The momentum transforms as the components of a covector, so it will change as the inverse of that constant: momentum also reflects. Therefore $p_j$ will become $-p_j$. The change during the evolution becomes $(-\frac{dx^i}{ds},\frac{dt}{ds},-\frac{dp_j}{ds},\frac{dE}{ds})$.

Suppose we perform a time reflection on top of the spatial reflection. This will send $t$ to $-t$. And since energy transforms as the time component of a covector, $E$ will become $-E$. The change during the evolution becomes $(-\frac{dx^i}{ds},-\frac{dt}{ds},-\frac{dp_j}{ds},-\frac{dE}{ds})$.

The change is now exactly the opposite on all components. The same change could have been achieved by changing the parametrization from $s$ to $-s$. But changing the sign of the parametrization means changing what we label as particle to what we label as anti-particle. If we call this our analogue for charge conjugation, we see that performing all three changes gets us back where we started.

\section{Conclusion}

Note that I would not considered an exact analogue for CPT. The other analogies we have found so far between classical/quantum/relativity are a lot stronger. Still, I think it points our intuition in the right direction and I don't think is just a coincidence either.

\end{document}

### 2018-02-13 Blog/18-PositionMomentum.tex
Adding blog posts

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Why are states identified by position and momentum?}

TL;DR - States are identified by position and momentum because this is the space that allows us to define coordinate-invariant densities.

As we saw in many of our posts, Hamiltonian mechanics describes the deterministic and reversible evolution of states. This leaves one question: why are states fully described by position and momentum? Why not only position? Why not position, momentum and something else? Why are state variables always in pairs? And why does momentum change in the opposite way during coordinate transformations?

The overall idea is the following: pairing variables in that way is the only way to define density distributions over the space of states (i.e. phase space) in a way that is invariant under coordinate changes. That is, if we want to have distributions of matter that are coordinate independent, such that the values are the same for all observers, then the distributions must be defined on pairs of conjugate quantities.

Let's see how this works.

\section{Distributions are not invariant}

Suppose you have a spatial distribution of mass $\rho(x^i)$. The unit of the distribution will be something like $kg/m^3$. Suppose we change units from $m$ to $km$. The density will be expressed in $kg/km^3$ therefore the value of our density will change accordingly by a factor of one billion. That is, the value of the density depends on the unit we use for space.

This presents a conceptual problem. We are typically used to thinking we have a density of mass at a point. But identifying the point and the unit used for mass is not enough to determine the density: we need to specify a unit for space as well. Why is this a conceptual problem? Suppose we have a composite object, like a finite amount of water, made of many pieces, modeled as infinitesimal parts of water. The state of the whole system will be a distribution over the states of the pieces. The fixed amount of water will be a distribution over the states of infinitesimal amounts of water (i.e. their position, momentum and whatever else). Now, if the distribution cannot even be defined without a coordinate system, how can this be an objective part of nature? Coordinate systems are things we create: nature does not know or care about them.

\section{Search for invariant densities}

Well, we know that the whole system must be distributed over the states of the infinitesimal parts. And we know such a distribution must be invariant under coordinate transformations. Both facts combined tell us that the state space of the infinitesimal parts cannot be arbitrary: it has to be such that if we change the unit for one variable, the density is not affected. Which means, the volume of the state space must be invariant as well.

The only way this can happen is if each variable that defines the coordinate system is paired with a variable that uses the inverted units. For example, if we have $x$ expressed in meters, then we must have $k_x$ expressed in inverse meters. Suppose we have $\Delta x = 1 \, m$ and $\Delta k_x = 1 \, \frac{1}{m}$, then $\Delta x \Delta k_x = 1$. Suppose we change units to kilometers, we have $\Delta x = 0.001 \, km$, $\Delta k_x = 1000 \, \frac{1}{km}$ and $\Delta x \Delta k_x = 1$. The volume does not change and therefore the density we define on it does not change either.

As we said, this is the only way to do it. If we had only $x$, then the density would not be invariant. If we had more than two variables, then simply knowing the units (or the change of units) of $x$ would not be enough to determine the units of the other variables. That is: the change of unit of $x$ and the invariance of the phase space volume provide two constraints so, if the system is well defined, we can only have two variables.

Note the phase space volume is in ``natural units": it's a pure number. We define $\hbar$ to be the unit of measurement for state space volume along one degree of freedom (i.e. a pair of conjugate variables). We can set $p_x = \hbar k_x$ so that the product $\Delta x \Delta p_x = \hbar \Delta x \Delta k_x$ already gives us the phase space volume with the correct units. And $p_x$ is exactly conjugate momentum.

\section{Conclusion}

Hamiltonian mechanics is the deterministic and reversible evolution of distributions that are defined independently of coordinates. The phase space for infinitesimal parts is charted by position and momentum because this is the space that allows us to define coordinate-invariant densities.

\end{document}

### 2018-02-13 Blog/19-ComplexNumbers.tex
Adding blog posts

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{What are complex numbers?}

TL;DR - Complex numbers should really be called rotation numbers. Whenever they are used in physics and engineering, some type of rotation is lurking and begging to be understood.

In the past posts we went through a lot of insights in classical mechanics and I want to start doing the same for quantum mechanics. But first we have to talk about complex numbers. Complex numbers are prominent in quantum mechanics and, before we understand why that is, we have to understand what complex numbers are in general.

Unfortunately, complex and imaginary numbers have names that are not very meaningful: they are not complex and they are not imaginary. The other problem is that they are typically introduced in the way they were discovered historically (i.e. solving polynomial equations) and we drag along this idea of $\imath = \sqrt{-1}$. None of this is intuitive or useful to understand why complex numbers are important in physics and engineering.

Briefly: complex numbers describe 2D vectors (i.e. a pair of real numbers) and rotations acting on them. You may sort of know this already. You may not realize, though: that's the only thing they do. Let's see how this work.

\section{Sets of numbers and bad naming}

While we are on the subject, note that complex numbers are not the only horribly named set of numbers in the English language. So, let's go through all of them.

First, we have the natural numbers $\mathbb{N} = \{0, 1, 2, ...\}$. Now, there is nothing ``natural" about them or ``unnatural" about the other numbers. These are the numbers we use to count: counting numbers would be a better name for $\mathbb{N}$.

Next we have the integer numbers $\mathbb{Z} = {..., -2, -1, 0, 1, 2, ...}$. The name comes from the same Latin root as ``entire". Whole numbers would be a better English name for $\mathbb{Z}$ (and sometimes they are called just that).

We can then take a pair of integer number, a numerator and a denominator, and construct the rational numbers $\mathbb{Q}$. They are not ``rational" in the sense of ``sensible". They are rational because they express a ratio, a proportion between two integer values. Same goes for irrationals: they are not crazy numbers, they simply cannot be expressed as a ratio. Fractional numbers would probably be a less confusing name for $\mathbb{Q}$.

And finally we have the real numbers $\mathbb{R}$ which are defined as the distinct Cauchy sequences of rational numbers. There is nothing ``real" about them. What they represent are values that in general can only be approximated by rationals. Physically, we use them when we assume we can measure values with arbitrary precision, so arbitrary precision numbers would be a more precise name for $\mathbb{R}$.

Now, despite that the names are not actually that good, you most likely have a good intuitive understanding of these types of numbers because they were introduced with real world examples. Hopefully.

\section{Rethinking complex numbers}

So how should we define and think about the complex numbers $\mathbb{C}$? As you know, a complex number $c$ is simply a pair of real numbers (i.e. arbitrary precision numbers). Instead of noting it $a+\imath b$, which makes the two elements look different, we can write it as $c = a\mathbf{1} + b \mathbf{i}$. Instead of calling $a$ the ``real part", let's call it the ``horizontal part". Instead of calling $b$ the ``imaginary part", let's call it the ``vertical part".

If we have two complex numbers $c_1$ and $c_2$, we can define the sum as $c_1+c_2 = (a_1+a_2) \mathbf{1} + (b_1+b_2)\mathbf{i}$: the horizontal part is the sum of the horizontal parts and the vertical part is the sum of the vertical parts. Note how the new name makes it seem beyond obvious: it's just a standard vector with two components. Note that $\mathbf{1}$ and $\mathbf{i}$ are simply the unit vectors along the horizontal and vertical directions. And the norm is given by $|c| = \sqrt{a^2 + b^2}$

Now suppose we want to characterize the linear operations. Let's first look at how they work in $\mathbb{R}$. Each number $a \in \mathbb{R}$ can be associated with the operation of rescaling by that number $f_a(x) = a x$. Each complex number $c \in \mathbb{C}$, instead, will be associated with a rescaling by the norm $|c|$ and a rotation by the angle between $c$ and the horizontal direction $\mathbf{1}$.

Since the operation is linear, we really just need to define the multiplication between the basis vectors. We have $\mathbf{1} \cdot \mathbf{1} = \mathbf{1}$ and $\mathbf{1} \cdot \mathbf{i} = \mathbf{i}$: since the angle between $\mathbf{1}$ and $\mathbf{1}$ is zero, the rotation leaves the direction unchanged. We also have $\mathbf{i} \cdot \mathbf{1} = \mathbf{i}$: since $\mathbf{i}$ is in the vertical direction, the corresponding rotation will move the horizontal by 90 degrees to the vertical direction. Lastly, $\mathbf{i} \cdot \mathbf{i} = - \mathbf{1}$: we are rotating the vertical direction by another 90 degrees so we obtain the opposite of the horizontal direction.

Note that we have obtained the usual relationship $\imath^2=-1$, but this time we know what we are talking about and there is nothing imaginary about it. It simply means that if we rotate 90 degrees twice, we get the opposite direction than we started with. It should also make it clear why we can express $c=\rho (\cos \theta \mathbf{1} + \sin \theta \mathbf{i})$. Mathematically, there shouldn't really be anything new about it but I hope that intuitively it clicked that there is nothing mysterious or arbitrary about any of this.

Complex numbers are just 2D vectors and rotations: rotation numbers would be a much less confusing name for $\mathbb{C}$.

\section{An example: linear systems}

Let's make complex numbers a little bit more concrete by quickly showing some highlights from linear control theory and signal processing. Suppose you have a system that takes an input signal $I(t)$ and converts it to an output signal $O(t)$. Suppose the system is linear: if we sum two inputs and feed them into the system we simply get the sum of the independent outputs. If that's the case, on can show that the relationship between inputs and outputs is of the form:
\begin{equation}
	O(t) = \mu_0 I(t) + \mu_1 \frac{d}{dt} I(t) + \mu_2 \frac{d^2}{dt^2} I(t) + ... + \mu_n \frac{d^n}{dt^n} I(t) + ...
\end{equation}

Now, suppose that the input is of the form $I(\omega, t) = a_\omega \cos \omega t$. The derivative will change cosines to sines and vice-versa, so the result will still be an oscillation of the form $O(t) = \hat{a}_\omega \cos \omega t + \hat{b}_\omega \sin \omega t$. We can set $\mathbf{1} = \cos \omega t$ and $\mathbf{i} = \sin \omega t$. So we have $I(\omega, t) = a_\omega \mathbf{1} = c_\omega$ and $O(\omega, t) = \hat{a}_\omega \mathbf{1} + \hat{b}_\omega \mathbf{i} = \hat{c}_\omega$. That is: we can express each frequency component with a two dimensional vector and the effect of the linear system is simply a rotation in that space. The linear system may change the magnitude (i.e. the strength of the signal) and the angle (i.e. the phase of the signal) but nothing else (i.e. not the shape of the signal).

Now, the point here is not to give a full introduction to control theory (which would take a lot more space). Hopefully, this should give you a sense of when and why rotation numbers, sorry, complex numbers are useful in physics and engineering: they will always be useful when we are studying the effect of linear systems, and all systems are, at first approximation, linear.

\section{Conclusion}

Complex numbers are not that complex to understand: they just describe 2D vectors and rotations. They are unfortunately named and the notation typically used is meaningless for the purpose of physics and engineering. It is important to understand that these are historical accidents, and they have no bearing on the actual concepts and their applications.

When we are solving complex polynomial equations, then, we are not looking for quantities: we looking for rotations that satisfy a particular property. For example, $x^2=-1$ is asking: what is the rotation that, if applied twice, is equivalent to a reflection? It's the 90 degree rotation, of course, which is denoted by $\imath$.

Complex numbers are fundamental when studying linear systems. Quantum mechanics, as we'll see, studies linear systems and that is why complex numbers are prevalent. You just need to understand what physical quantity is the 2D vector representing and what is rotating.

\end{document}

### 2018-02-13 Blog/20-NotAllObservablesAreOperators.tex
Adding blog posts

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Not everything we measure is an eigenvalue of a linear operator}

TL;DR - Statistical quantities (e.g. averages) and angles (e.g. direction of spin) are measurable quantities but are not associated with linear operators, eigenkets and eigenvalues.

When studying quantum mechanics you learn about observables, how to each you associate a Hermitian operator, how the value is only defined on the eigenstates of that operator and how, in general, you will have a distribution over eigenvalues. Position, momentum, energy and spin are all examples. Since one mostly deals with those, one usually gets the impression that that's all there is. This may not be stated per se in your textbook, yet you may have that impression.

But is that true? Is everything that we measure an eigenvalue of some Hermitian operator? Here I'll present two quantities that don't follow the pattern: temperature and direction of spin.

\section{Temperature}

Suppose we have a box filled with gas in thermodynamic equilibrium. Its temperature $T$ will be proportional to the variance of the velocity of all the elementary constituents of the gas. If we call $|\psi>$ the state of all the particles, $P_i$ the momentum operator for the $i$-th particle and $m_i$ its mass, we'll have something like:
\begin{equation}
T=\alpha<\psi|\sum_{i=1}^{n} \frac{P_i^2}{m_i}|\psi>
\end{equation}
where $\alpha$ is an appropriate constant.

Now, what's important here is not the detail of the expression: the important aspect is that temperature is an average. Any state that represents a snapshot of a system (i.e. a pure state) will always have one and only one value of temperature. And that value needs to match what our thermometer says. That is: we are not going to have a statistical distribution over possible values of temperatures.

You may think: but the quantity $\sum_{i=1}^{n} \frac{P_i^2}{m_i}$ is an operator. And indeed it is. But that's where the connection with temperature ends. Think about the eigenstates of that operator: they correspond to those states for which the magnitude of the momentum is perfectly prepared for all particles. Those are not the only states for which we have a well defined value of temperature. And measuring the temperature does not mean measuring the magnitude of the momentum for each particle. So $\sum_{i=0}^{n} \frac{P_i^2}{m_i}$ is an operator but is not the temperature operator: it's an operator whose expectation corresponds to the temperature.

\section{Spin direction}

Now suppose we have a spin 1/2 system. You may be familiar with $S_x$, $S_y$ and $S_z$ which are the operators for the spin components along the respective directions. As you know, spin represents angular momentum so their conjugate quantity is the angle along the plane perpendicular to their direction. Which leads to the question: where is the spin angle operator? Well, there isn't one.

All spin 1/2 states can be defined by a unique direction in space. We can write:
\begin{equation}
|\psi>=\cos(\theta/2)|z^+> + \sin(\theta/2)e^{\imath \phi}|z^->
\end{equation}
where $\theta$ and $\phi$ are the polar and azimuthal angle respectively. Note that we just need two states, $|z^+>$ and $|z^->$, to form a basis and those correspond to the two possible values of spin measured along the $z$ direction. An angle, instead, takes a continuum of possible values and therefore we would need an infinite number of eigenkets to form an angle operator (as it is for position and momentum). Since the space is two dimensional, all bases must be two dimensional: no angle operator.

But here is the thing: we can nonetheless measure angles. Suppose we have a source of electrons such that their spin comes aligned always in the same direction. With a Stern-–Gerlach type experiment, we can measure the fraction $0\leq f_z \leq 1$ that comes out with $z^+$. We have $f_z = <\psi|z^+><z^+|\psi> = \cos^2(\theta/2)$. So $\theta = 2 \arccos \sqrt{f_z}$ is definitely something we can measure. Similarly, we can find an expression for $\phi$.

Again you may think: all we did was measure the expectation of $|z^+><z^+|$ which is a Hermitian operator. And indeed we did. But that operator has eigenstates only for $\theta=0$ and $\theta=\pi$. Yet each state will always have a well defined value for $\theta$ and we can measure the values between $0$ and $\pi$ as well.

\section{Conclusion}

While it is often useful to think of a quantum state as a distribution over the eigenvalues of some observable, this is not the only way we should think about it. Not all measurable quantities work like that. In particular, note that many macroscopic quantities are averages over a large number of particles and therefore one should always be very careful when extrapolating ideas from the quantum world.

\end{document}

### 2018-02-13 Blog/21-WhatAreCommutators.tex
Adding blog posts

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{What are commutators?}

TL;DR Commutators, like Poisson bracket, tells how a quantity changes under a transformation generated by another.

Commutators play a fundamental role in quantum mechanics. What we want to show is what they are in relation to infinitesimal transformation, which gives a better understanding of why they are related to Poisson brackets and why they must give the relationships they give.

\section{Conjugate quantities}

In classical mechanics, conjugate quantities are pair of variables that form an independent degree of freedom. If you integrate over a pair, you get a quantity that no longer depends on the unit and coordinate used for those quantities.

In quantum mechanics, things work a bit differently mathematically. In quantum mechanics you have operators and parameters. If a quantity is an operator it means you can write the state as a wave-function over the quantity. If a quantity is an operator it means you can make and infinitesimal transformation along it. The conjugate relationship is between operators and parameters: for each operator $A$ you have one parameter $\alpha$ and we say that $A$ generates the infinitesimal transformation $(1 + \frac{A d\alpha}{\imath \hbar})$, which is a unitary transformation.

For example, spin component $S_z$ is the operator while the angle $\theta_{xy}$ is the parameter, and $S_z$ generates the rotation $(1 + \frac{S_z d\theta_{xy}}{\imath \hbar})$. The confusing part is that some quantities, like position and momentum, can both be operators and parameters. So we have the $P$ operator and the $x$ parameter and $P$ generates the translation $(1 + \frac{P dx}{\imath \hbar})$, but also $X$ operator and $p$ parameter and $X$ generates the change in momentum $(1 + \frac{X dp}{\imath \hbar})$.

\section{Commutators}

Now we ask: suppose we have two operators $A$ and $B$. Suppose you perform the infinitesimal transformation generated by the second, over the parameter $\beta$. How does the first change?

The infinitesimal transformation corresponds to the unitary operator $U=(1 + \frac{B d\beta}{\imath \hbar})$. An operator under a unitary transformation changes as $\hat{A} = U^T A U$. Therefore we have:

\begin{equation}
\begin{aligned}
\hat{A}&=U^\dagger A U = (1 - \frac{B d\beta}{\imath \hbar}) A (1 + \frac{B d\beta}{\imath \hbar}) \\
&=A + A \frac{B d\beta}{\imath \hbar} - \frac{B d\beta}{\imath \hbar} A - \frac{B d\beta}{\imath \hbar} \frac{B d\beta}{\imath \hbar} \\
&=A + \frac{(AB - BA)}{\imath \hbar}d\beta + \frac{B^2}{\hbar^2}d\beta^2
\end{aligned}
\end{equation}

The first order change of $A$ is therefore:

\begin{equation}
\begin{aligned}
\frac{dA}{d\beta}&=\frac{(AB - BA)}{\imath \hbar} \\
&= \frac{[A,B]}{\imath \hbar}
\end{aligned}
\end{equation}

Therefore the commutator between $A$ and $B$ tells us how $A$ changes under a transformation generated by $B$. In a previous post we saw that the Poisson brackets in classical Hamiltonian mechanics had the same role. That is why we can do the formal substitution: they describe the same physical relationship in the different mathematical framework.

We can also understand why the commutators must be what they are. For example, the conjugate of the operator momentum $P_j$ is the translation over the position parameter $x_j$. So how will the operator position $X_i$ change under the transformation generated by $P_j$?

\begin{equation}
\begin{aligned}
\frac{[X_i,P_j]}{\imath \hbar} &= \frac{dX_i}{dx_j}=I \delta_{ij}
\end{aligned}
\end{equation}

If $X_i$ and $P_j$ are along the same direction, then position operator will change of the same amount of the position parameter: they are physically the same thing. If they are along different directions there is no change.

Now, consider the operators $S_x$, $S_y$ and $S_z$ for the spin components. How will they change under the transformation generated by $S_z$, which is the rotation over $\theta_{xy}$?

\begin{equation}
\begin{aligned}
\frac{[S_x,S_z]}{\imath \hbar} &= \frac{dS_x}{d\theta_{xy}} = S_y \\
\frac{[S_y,S_z]}{\imath \hbar} &= \frac{dS_x}{d\theta_{xy}} = - S_x \\
\frac{[S_z,S_z]}{\imath \hbar} &= \frac{dS_z}{d\theta_{xy}} = 0
\end{aligned}
\end{equation}

The last one is the easiest: the $z$ component of spin does not change under rotation over the $xy$ plane. For the other two, just remember that the $x$ component is moved in the $y$ direction and the $y$ direction is moved in the $-y$ direction.

These are indeed the commutation relationship one has in quantum mechanics. Given the physical meaning of the operators and their transformations, the commutation relationship can't be anything else.

\section{Conclusion}

We have seen that commutators have a well defined physical/geometrical meaning, which is the same one for Poisson brackets in classical Hamiltonian mechanics. I personally not only find it insightful, but it helps me remember the correct sign for the commutation relationships.

\end{document}

### 2018-02-13 Articles/TopologyAndExperimentalDistinguishability/article-topologyproceedings.tex
Version of the article submitted to topology proceedings

--- 2017-11-25 Articles/TopologyAndExperimentalDistinguishability/article.tex
@@
-\documentclass[preprint]{elsarticle}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%           Topology Proceedings Sample Latex Article             %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% Temporary to remove "submitted to"
-\makeatletter
-\def\ps@pprintTitle{%
-	\let\@oddhead\@empty
-	\let\@evenhead\@empty
-	\def\@oddfoot{}%
-	\let\@evenfoot\@oddfoot}
-\makeatother
+\documentclass{amsart}
+\usepackage[T1]{fontenc}   %%%%%%%Please do not change
+%\usepackage{graphics}                  %%%%%%%
 
-\usepackage{lineno,hyperref}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Please do not change this paragraph. %%%%%%%
+\setcounter{page}{1}                   %%%%%%%
+\setlength{\textwidth}{4.4in}          %%%%%%%
+\setlength{\textheight}{7.0in}         %%%%%%%
+\setlength{\evensidemargin}{1in}       %%%%%%%
+\setlength{\oddsidemargin}{1in}        %%%%%%%
+\setlength{\topmargin}{.8in}           %%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Please do not change the page size   %%%%%%%
+% and do not redefine the baselineskip.%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\usepackage{amsmath, amsthm, amsfonts}
-\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}
-
-\newcommand{\R}{\mathbb{R}}
-\newcommand{\Q}{\mathbb{Q}}
-\newcommand{\Z}{\mathbb{Z}}
-
-\theoremstyle{plain}% default 
 \newtheorem{thm}{Theorem}[section]
@@
 \newtheorem{prop}[thm]{Proposition}
-\newtheorem{cor}[thm]{Corollary}
-
 \theoremstyle{definition}
 \newtheorem{defn}[thm]{Definition}
-%\newtheorem{exmp}{Example}[section]
-\theoremstyle{remark}
-\newtheorem*{rem}{Remark}
-\modulolinenumbers[5]
 
-\journal{Topology and its Applications}
+\newtheorem{rem}[thm]{Remark}
+\numberwithin{equation}{section}
+
+\usepackage{amsmath, amsthm, amsfonts, float}
+\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}
 
@@
 
-\begin{frontmatter}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% This a placeholder for the TOPLOGY PROCEEDINGS logo %%%%%%
+\noindent                                             %%%%%%
+\begin{picture}(150,36)                               %%%%%%
+\put(5,20){\tiny{Submitted to}}                       %%%%%%
+\put(5,7){\textbf{Topology Proceedings}}              %%%%%%
+\put(0,0){\framebox(140,34){}}                        %%%%%%
+\put(2,2){\framebox(136,30){}}                        %%%%%%
+\end{picture}                                        %%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{0.5in}
+
+\renewcommand{\bf}{\bfseries}
+\renewcommand{\sc}{\scshape}
+%insert defs/styles
+\vspace{0.5in}
 
@@
 
-%% Group authors per affiliation:
+%    Information for first author:
+\author{Christine A. Aidala}
+\address{Department of Physics; University of Michigan; Ann Arbor, MI 48109}
+%    Current address (if needed):
+%\curraddr{}
+\email{caidala@umich.edu}
+%\thanks{The first author was supported in part by NSF Grant \#000000.}
 
-\author[]{Christine A. Aidala}
-\ead{caidala@umich.edu}
+\author{Gabriele Carcassi}
+\address{Department of Physics; University of Michigan; Ann Arbor, MI 48109}
+\email{carcassi@umich.edu}
 
-\author[]{Gabriele Carcassi\corref{cor1}}
-\ead{carcassi@umich.edu}
+\author{Mark J. Greenfield}
+\address{Department of Mathematics; University of Michigan; Ann Arbor, MI 48109}
+\email{markjg@umich.edu}
 
-\author[]{Mark J. Greenfield}
-\ead{markjg@umich.edu}
+%    General info
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\subjclass[2010]{54H99, 00A79}
+%                                                                                                                           %
+%         Please use the current 2010 Mathematics Subject Classification:             %
+%         http://www.ams.org/mathscinet/msc/                                                        %
+%         http://www.zentralblatt-math.org/msc/en/                                                 %
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-\address{University of Michigan, Ann Arbor, MI 48109, USA}
-
-\cortext[cor1]{Corresponding author}
+\keywords{topological spaces, scientific foundations, function space}
+\thanks {Funding for this work was provided in part by the MCubed program of the University of Michigan.}
 
@@
 
-\begin{keyword}
-topological spaces \sep scientific foundations \sep function space
-\MSC[2010] 54H99 \sep 00A79
-\end{keyword}
+\maketitle
 
-\end{frontmatter}
-
-%\linenumbers
-
-\section{Introduction}
+\section{\bf Introduction}
 
@@
 
-\section{Experimental observations}
+\section{\bf Experimental observations}
 
@@
 
-\section{Algebra of experimental observations}
+\section{\bf Algebra of experimental observations}
 
@@
 
-\section{Experimental domain}
+\section{\bf Experimental domain}
 
@@
 
-\section{Experimental distinguishability}
+\section{\bf Experimental distinguishability}
 
@@
 
-\section{Experimental relationships}
+\section{\bf Experimental relationships}
 
@@
 
-\section{Distinguishability of experimental relationships}
+\section{\bf Distinguishability of experimental relationships}
 
@@
 
-\begin{table}[h]
-	\centering
-	\begin{tabular}{p{0.20\textwidth} p{0.7\textwidth}}
-		Math/Topology & Science/Physics \\ 
-		\hline 
-		Hausdorff, second-countable space & Space of experimentally distinguishable elements, whose points are the possible values and whose open sets represent the experimentally attainable levels of precision.
-\\
-		Open set & Verifiable set. We can verify experimentally that an element is within the set.  \\ 
-		Closed set & Refutable set. We can verify experimentally that an element is not in the set. \\ 
-		Basis & A collection of verifiable sets that can be used to distinguish an element\\
-		Continuous \newline function &  An experimental relationship between two sets of experimentally distinguishable elements, which must preserve distinguishability \\
-		Homeomorphism &  A perfect equivalence between spaces of experimentally distinguishable elements. \\
-	\end{tabular} 
-	\caption{Topology-to-physics dictionary. This table sums up the relationships established between mathematical and scientific concepts.}
-\end{table}
-
-\section{Conclusion}
+\section{\bf Conclusion}
 
@@
 
-\section*{Acknowledgments}
-Funding for this work was provided in part by the MCubed program of the University of Michigan.
+\begin{table}[H]
+	\centering
+	\begin{tabular}{p{0.20\textwidth} p{0.7\textwidth}}
+		Math/Topology & Science/Physics \\ 
+		\hline 
+		Hausdorff, 2nd-countable space & Space of experimentally distinguishable elements, whose points are the possible values and whose open sets represent the experimentally attainable levels of precision.
+\\
+		Open set & Verifiable set. We can verify experimentally that an element is within the set.  \\ 
+		Closed set & Refutable set. We can verify experimentally that an element is not in the set. \\ 
+		Basis & A collection of verifiable sets that can be used to distinguish an element\\
+		Continuous \newline function &  An experimental relationship between two sets of experimentally distinguishable elements, which must preserve distinguishability \\
+		Homeo-morphism &  A perfect equivalence between spaces of experimentally distinguishable elements. \\
+		\\
+	\end{tabular} 
+	\caption{Topology-to-physics dictionary. This table sums up the relationships established between mathematical and scientific concepts.}
+\end{table}
 
-\bibliographystyle{elsarticle-num}
-\bibliography{bibliography}
+\bibliographystyle{plain}
+\begin{thebibliography}{10}
+
+	\smallskip
+
+	\bibitem{wigner} E. Wigner, {\it The unreasonable effectiveness of mathematics in the natural sciences}, Communications on Pure and Applied Mathematics 13 (1) (1960) 1-14. doi:10.1002/cpa.3160130102.
+
+	\smallskip
+
+	\bibitem{rashevsky} N.
+Rashevsky, {\it Topology and life: In search of general mathematical principles in biology and sociology}, The bulletin of mathematical biophysics 16 (4) (1954) 317-348.
+doi:10.1007/BF02484495.
+
+	\smallskip
+
+	\bibitem{chichilnisky} G.
+Chichilnisk, {\it Social choice and the topology of spaces of preferences}.
+Advances in Mathematics 37 (2) (1980) 165-176.
+doi:http://dx.doi.org/10.1016/0001-8708(80)90032-8.
+
+	\smallskip
+
+	\bibitem{nakahara} M. Nakahara, {\it Geometry, topology and physics}, CRC Press, 2003.
+
+	\smallskip
+
+	\bibitem{basener} W. F. Basener, eds. {\it Topology and its applications}, John Wiley \& Sons, 2006.
+
+	\smallskip
+
+	\bibitem{mill}  J. S. Mill, {\it A system of logic}, Parker, 1843
+
+	\smallskip
+
+	\bibitem{russell} B. Russell, {\it The problems of philosophy}, Home University Library, 1912.
+\end{thebibliography}
 

### 2018-02-15 Blog/21-WhatAreCommutators.tex
Minor edits

@@
 \section{What are commutators?}
-
-TL;DR Commutators, like Poisson bracket, tells how a quantity changes under a transformation generated by another.
+TL;DR - Commutators, like Poisson bracket, tell us how a quantity changes under a transformation generated by another.
 
 Commutators play a fundamental role in quantum mechanics.
-What we want to show is what they are in relation to infinitesimal transformation, which gives a better understanding of why they are related to Poisson brackets and why they must give the relationships they give.
+Mathematically, they tell us how multiplication between operators behaves but this does not provide any physical insight.
+What is more interesting physically is that they are intimately related to infinitesimal transformations, which gives a better understanding of why they are related to Poisson brackets and why commutation relationships between position, momentum and spin must be what they are.
 
@@
 In quantum mechanics you have operators and parameters.
-If a quantity is an operator it means you can write the state as a wave-function over the quantity.
-If a quantity is an operator it means you can make and infinitesimal transformation along it.
+If a quantity is an operator (like spin) it means you can write the state as a wave-function over that quantity.
+If a quantity is a parameter (like angle) it means you can make and infinitesimal transformation along it.
 The conjugate relationship is between operators and parameters: for each operator $A$ you have one parameter $\alpha$ and we say that $A$ generates the infinitesimal transformation $(1 + \frac{A d\alpha}{\imath \hbar})$, which is a unitary transformation.
@@
 So we have the $P$ operator and the $x$ parameter and $P$ generates the translation $(1 + \frac{P dx}{\imath \hbar})$, but also $X$ operator and $p$ parameter and $X$ generates the change in momentum $(1 + \frac{X dp}{\imath \hbar})$.
+Note how the upper case denotes an Hermitian operator while the lower case denotes a real number.
 
@@
 
-Now we ask: suppose we have two operators $A$ and $B$. Suppose you perform the infinitesimal transformation generated by the second, over the parameter $\beta$. How does the first change?
+Now we ask: suppose we have two operators $A$ and $B$. Suppose you perform the infinitesimal transformation generated by the second, over the parameter $\beta$. How does the operator $A$ change?
 
 The infinitesimal transformation corresponds to the unitary operator $U=(1 + \frac{B d\beta}{\imath \hbar})$.
-An operator under a unitary transformation changes as $\hat{A} = U^T A U$.
+An operator under a unitary transformation changes as $\hat{A} = U^\dagger A U$.
 Therefore we have:
@@
 \begin{aligned}
-\hat{A}&=U^\dagger A U = (1 - \frac{B d\beta}{\imath \hbar}) A (1 + \frac{B d\beta}{\imath \hbar}) \\
-&=A + A \frac{B d\beta}{\imath \hbar} - \frac{B d\beta}{\imath \hbar} A - \frac{B d\beta}{\imath \hbar} \frac{B d\beta}{\imath \hbar} \\
-&=A + \frac{(AB - BA)}{\imath \hbar}d\beta + \frac{B^2}{\hbar^2}d\beta^2
+\hat{A}&amp;=U^\dagger A U = (1 - \frac{B d\beta}{\imath \hbar}) A (1 + \frac{B d\beta}{\imath \hbar}) \\
+&amp;=A + A \frac{B d\beta}{\imath \hbar} - \frac{B d\beta}{\imath \hbar} A - \frac{B d\beta}{\imath \hbar} A \frac{B d\beta}{\imath \hbar} \\
+&amp;=A + \frac{(AB - BA)}{\imath \hbar}d\beta + \frac{BAB}{\hbar^2}d\beta^2
 \end{aligned}
@@
 
-The first order change of $A$ is therefore:
+The first order for the change of $A$ is:
 
@@
 \begin{aligned}
-\frac{dA}{d\beta}&=\frac{(AB - BA)}{\imath \hbar} \\
-&= \frac{[A,B]}{\imath \hbar}
+\frac{dA}{d\beta}&amp;=\frac{(AB - BA)}{\imath \hbar} \\
+&amp;= \frac{[A,B]}{\imath \hbar}
 \end{aligned}
@@
 In a previous post we saw that the Poisson brackets in classical Hamiltonian mechanics had the same role.
-That is why we can do the formal substitution: they describe the same physical relationship in the different mathematical framework.
+That is why we can do the formal substitution: they describe the same physical relationship in a different mathematical framework.
 
-We can also understand why the commutators must be what they are.
-For example, the conjugate of the operator momentum $P_j$ is the translation over the position parameter $x_j$.
+We can also understand why commutators between notable quantities must be what they are.
+For example, the infinitesimal transformation generated by the operator momentum $P_j$ is the translation over the position parameter $x_j$.
 So how will the operator position $X_i$ change under the transformation generated by $P_j$?
@@
 \begin{aligned}
-\frac{[X_i,P_j]}{\imath \hbar} &= \frac{dX_i}{dx_j}=I \delta_{ij}
+\frac{[X_i,P_j]}{\imath \hbar} &amp;= \frac{dX_i}{dx_j}=I \delta_{ij}
 \end{aligned}
@@
 \begin{aligned}
-\frac{[S_x,S_z]}{\imath \hbar} &= \frac{dS_x}{d\theta_{xy}} = S_y \\
-\frac{[S_y,S_z]}{\imath \hbar} &= \frac{dS_x}{d\theta_{xy}} = - S_x \\
-\frac{[S_z,S_z]}{\imath \hbar} &= \frac{dS_z}{d\theta_{xy}} = 0
+\frac{[S_x,S_z]}{\imath \hbar} &amp;= \frac{dS_x}{d\theta_{xy}} = S_y \\
+\frac{[S_y,S_z]}{\imath \hbar} &amp;= \frac{dS_x}{d\theta_{xy}} = - S_x \\
+\frac{[S_z,S_z]}{\imath \hbar} &amp;= \frac{dS_z}{d\theta_{xy}} = 0
 \end{aligned}
@@
 The last one is the easiest: the $z$ component of spin does not change under rotation over the $xy$ plane.
-For the other two, just remember that the $x$ component is moved in the $y$ direction and the $y$ direction is moved in the $-y$ direction.
+For the other two, just remember that the $x$ component is rotated in the $y$ direction and the $y$ direction is rotated in the $-x$ direction.
 
-These are indeed the commutation relationship one has in quantum mechanics. Given the physical meaning of the operators and their transformations, the commutation relationship can't be anything else.
+These are indeed the commutation relationships one has in quantum mechanics. Given the physical meaning of the operators and their transformations, the commutation relationship can't be anything else.
 

### 2018-04-25 Blog/22-AllQuantumStatesAreEquallyUncertain.tex
Adding blog posts

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{All quantum states are equally defined}
TL;DR - No quantum state is more uncertain than the other. All states can be identified by a set of perfectly prepared quantities.

When studying quantum mechanics, you may get the (wrong) impression that some states are better defined than others. Some are eigenstates while others are just a superposition. Or that the gaussian packets, are more determined than the other states, since some satisfy the uncertainty principle with an equal. Well, that's just not the case.

The confusion stems from the idea that quantum states are like classical statistical distribution, in which you have some well defined elements (which are the objects that are well defined) upon which you assign probabilities. Quantum states are not like this at all. We are going to see that any state is always the eigenstate of some Hermitian operator, and therefore they always have a quantity that is perfectly prepared. And they also have a symmetry under the transformation generated by that operator, so there is always another quantity of which no value is more likely than the other.

\section{Prepared quantities and unprepared conjugates}

Suppose that you have a quantum state for which the position is perfectly prepared. This corresponds to the eigenstate $|x\rangle$ of the operator $X$. The transformation generated by $X$ is $1+\frac{Xdp}{\imath\hbar}$ which corresponds to increasing momentum by $dp$. Now, since $|x\rangle$ is an eigenstate of $X$, it is also an eigenstate of the transformation generated by $X$. If we imagine the distribution of $|x\rangle$ over momentum, then, this has to be a distribution that does not change if we increase momentum. But the only distribution that is symmetric under that change is the one that has the same value for all possible values of momentum: the eigen state $|x\rangle$ is uniformly distributed in momentum. This can be generalized to any quantity. For example, an eigenstate of spin in the $z$ direction will be symmetric along the angle on the $(x,y)$ plane. This will also hold for any function of position and momentum and indeed for any Hermitian operator.

So, for a particular space of quantum states, we can imagine that, instead of specifying a wavefunction, we can give a set of operators and eigenvalues. Given that information, we can identify the corresponding eigenstate. This would not give the value for all quantities since the conjugate quantities will be left completely unspecified. The question is: can all states can be specified in this way? Which is equivalent to asking: are all states eigenstates of some Hermitian operator?

Suppose we have a state $|\psi\rangle$. We can construct the operator $O=|\psi\rangle a \langle \psi |$ where $a$ is a real number. $O$ is Hermitian and $|\psi\rangle$ is the eigenstate corresponding to the eigenvalue $a$. Yes, the operator is trivial, but we can indeed construct it. Which shows that for any state there exists Hermitian operators that allow that state as an eigenstate. In fact, there are infinitely many.

Overall, a state always has some well specified quantities and it also has some unspecified ones. For example, suppose we have an eigenstate for the operator $X + a P$, a linear combination of position and momentum. This means that the quantity $x + ap$ is perfectly prepared while the conjugate quantity, $\frac{1}{2a}(x-ap)$, is a uniform distribution. We can verify the two quantities are conjugates by calculating the commutator.

\begin{equation}
\begin{aligned}
\frac{\left[\frac{1}{2a}(X-aP), X+aP \right]}{\imath \hbar} &= \frac{1}{2a \imath \hbar} \left[X-aP, X+aP\right] \\
 &= \frac{1}{2a \imath \hbar} \left( \left[X, X\right] + \left[-aP, X\right] + \left[X, aP\right] + \left[-aP, aP\right] \right) \\
&= \frac{1}{2a \imath \hbar} \left( 0 + a \imath \hbar + a \imath \hbar + 0 \right) \\
&= 1
\end{aligned}
\end{equation}

Now consider the two distributions for that state over position and momentum. Since all the values of $\frac{1}{2a}(x-ap)$ are equally likely, then all values of $x$ and $p$ are also equally likely. That is: the distribution is uniform for both quantities. The variance for both $x$ and $p$ is infinite. If we only know that, we would think that state to be infinitely less defined than an eigetnstate of position. What happens is that for that state the uniform distribution in $x$ and $p$ are strongly correlated, so much so that the distribution over $x+ap$ admits a single value. Therefore this state is no less defined of an eigenestate of $x$: what changes is what quantities are well defined and which aren't.

\section{Conclusion}

What is significantly different in quantum mechanics is that all states are distributions but no distribution is more defined than the other. Even if a state has an extremely large spread in a pair of conjugate variables, it will have a quantity somewhere that will be perfectly defined. There is no base that is mathematically better then the other: all states can be part of a basis and all states are superpositions in some other basis. Whenever we are superposing two states, we are always adding some other correlation such that the new states is as well defined as the other two.

This actually makes a lot of sense conceptually. It's telling us that each quantum state is a distribution that cannot be decomposed into smaller independent ones. Each quantum system is an irreducible unit. This is the main difference from classical mechanics, where each distribution over phase space can be divided into smaller pieces that can be studied independently.

\end{document}

### 2018-04-25 Blog/23-DeterminismAndQuantumMechanics.tex
Adding blog posts

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Determinism and quantum mechanics}
TL;DR - The Schrodinger equation can be seen as the deterministic and reversible limit of the projection associated with a measurement.

Quantum mechanics comes with two ways to go from an initial to a final state. The first is the Schrodinger equation that represents time evolution and the second is the projection that represents what happens during measurements. These are usually presented as two separate entities. In fact, much of the work surrounding various ``interpretations" is to reconcile these two types of evolution.

What we want to show here is that there is a very natural way to reconcile them. As we saw in a previous post, a quantum state always has a set of quantities that are well defined (i.e. it is always an eigenstate of some Hermitian operator). Therefore we can regard the Schrodinger equation as a special case of the projection where at each instant a measurement is made. Let's see how this works.

\section{The projection postulate}

Let's first review the projection postulate which supposedly describes what happens during measurements. The idea is that you start with a single well defined initial state $|\psi\rangle$. You pick an observable $A$, which is defined by a set of eigenstates $|\psi_{a_i}\rangle$ and the corresponding eigenvalues $a_i$. After the measurement you end up with what is called a mixed state: a statistical distribution. More precisely, a distribution over the eigenstates, each having probability $|\langle \psi_{a_i} | \psi \rangle | ^2$. So, for example, if we start with spin up $|s^+_z\rangle$ and we measure the horizontal direction, we end up with 50\% $|s^+_x\rangle$ and 50\% $|s^-_x\rangle$.

Note that we don't always end up with a mixed state: if the initial state is already an eigenstate of the observable, nothing changes. That is, if we start with spin up $|s^+_z\rangle$ and we measure the vertical direction, we end up with 100\% $|s^+_z\rangle$ and 0\% $|s^-_z\rangle$. Which is the same as what we started with. So projection postulate is not always non-deterministic. In fact we can go continuously from a process that is deterministic, where the direction of measurement is the same as the prepared direction, to one that is completely non-deterministic, where the direction of measurement is perpendicular to the prepared direction.

\section{Deterministic projections}

So the idea is the following: can we perform a measurement such that we are not exactly measuring the same observable (e.g. the same direction of spin) but something so close that the projection is still deterministic (e.g. the direction of spin in an infinitesimally close direction)? That is, we want the final state $|\psi_{t+dt}\rangle$ to be very close to the initial state $|\psi_t\rangle$ so that $|\langle \psi_{t+dt}|\psi_t\rangle|^2=|\langle \psi_t|\psi_{t+dt}\rangle|^2=1$. We can rewrite $|\psi_{t+dt}\rangle = (1 + dt \frac{\partial}{\partial t}) |\psi_t\rangle $. We have:

\begin{align*}
|\langle \psi_t|\psi_{t+dt}\rangle|^2 &= 1 = |\langle \psi_t | (1 + dt \partial_t) |\psi_t\rangle|^2 \\
&= \langle \psi_t | (1 + dt \partial_t)^\dagger |\psi_t\rangle \langle \psi_t | (1 + dt \partial_t) |\psi_t\rangle \\
&= (\langle \psi_t | \psi_t\rangle + \langle \psi_t | dt \partial_t^\dagger |\psi_t\rangle ) (\langle \psi_t | \psi_t\rangle + \langle \psi_t | dt \partial_t |\psi_t\rangle ) \\
&= 1 + dt (\langle \psi_t | \partial_t^\dagger |\psi_t\rangle + \langle \psi_t | \partial_t |\psi_t\rangle) + dt^2 (|\langle \psi_t | \partial_t |\psi_t\rangle|^2) \\
&\approx 1 + dt \langle \psi_t | \partial_t^\dagger + \partial_t |\psi_t\rangle \\
\partial_t &= - \partial_t^\dagger 
\end{align*}

We have that deterministic time evolution must be unitary and that its generator $\partial_t$ must be anti-Hermitian. This means we can find a corresponding Hermitian operator $H$ such that $\partial_t = \frac{H}{\imath \hbar}$. This gives us the Schrodinger equation.
\begin{align*}
	\frac{\partial}{\partial t} | \psi_t \rangle = \frac{H}{\imath \hbar} |\psi_t\rangle
\end{align*}

To put this in more concrete term, consider spin precession in a magnetic field. We can say that the magnetic field continuously measures the spin of the particle, and this is what causes the motion. The direction where this measurement happens is a function of the particle itself, which makes sense because the motion is deterministic: it only depends on the state of the particle.

\section{Conclusion}

We have seen that the Schrodinger equation is a continuous process that at every moment is measuring something that is very close to what was measured before. This way there is no difference between measurement and evolution between measurement: it is the same type of process. It also means that there is nothing special about a measurement: whether we cause it or not, is just the same process that happens all the time.

What this also means is that the deterministic and reversible process is the particular case, not the general one. But this is true anyway: it is only when we are able to isolate a system enough from the environment that we can assume that its evolution is independent from it.

Mathematically, we are associating continuous unitary evolution with deterministic and reversible processes while non-deterministic processes have no such restriction. This is very interesting because it is the same conclusion that we reach from a completely different premise in our work on the assumption of physics.

\end{document}

### 2018-04-25 Blog/24-QuantumUncertaintyNotCausedByMeasurements.tex
Adding blog posts

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-1}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Quantum uncertainty is not caused by measurements}
TL;DR - The uncertainty relationship in quantum mechanics exists before and independently of measurements.

Some books still perpetuate the idea that the uncertainty introduced by quantum mechanics is related to measurement. Either as an observer effect (i.e. the measurement causes the uncertainty) or as an observed relationship (i.e. our measurements are limited by that relationships). Unfortunately, neither of these are correct: the uncertainty exists before and independently of any measurement.

The problem is that in the early days physicists used these incomplete ideas as a stepping stone (e.g. Heisenberg's microscope). This is not uncommon in physics: Galileo himself argued that the tides of the oceans are clear indication that the earth is spinning. The difference is that nobody repeats Galileo's argument (though we keep the conclusion) while some early thought experiments in quantum mechanics are offered as insightful while they are actually misleading.

So let's look at the math and see what it tells us.

\section{The measurement collapse and the uncertainty}

Suppose you have an incoming electron, whose state is described by a wave function $\psi(x)$. Suppose we measure its position. According to the theory, the wave function will collapse in one of the eigenstates of position. That is, we have one wave function coming in, we have a distribution of wave function coming out. The probability associated with the eigenstate at a particular $x$ will be given by $\rho(x) = |\psi(x)|^2$. This distribution will have an associated standard deviation $\sigma_x = \sqrt{ \int x^2 \rho dx - (\int x \rho dx)^2}$.

Note that, since the outgoing states are all position eigenstates, the distribution in momentum for all of them is uniform. So the measurement has indeed introduced uncertainty. But all the uncertainty is introduced to momentum, the conjugate of position. The standard deviation $\sigma_x$ is the same because the distribution in position $|\psi(x)|^2$ is the same distribution we had in the wave function before the collapse. That is: the distribution over the observable we measure remains the same. If it didn't, in fact, we couldn't measure anything.

It's not that we can write wave functions for which the uncertainty principle is not satisfied, we collapse them, and then the uncertainty principle is found. Not at all. It's that we can't even write wave functions that violate the uncertainty principle. Let me repeat it once more for effect: you can't write a $\psi(x)$ for which the Fourier transform has an uncertainty on momentum that violates the uncertainty principle. It doesn't matter whether you are going to collapse it or not.

\section{Conclusion}

It's not that the uncertainties of our measurements are bound by the uncertainty principle. It's that we can't even prepare states that violate the uncertainty principle. Those states are excluded a priori: they do not exist in the theory.

Therefore we can't consider the uncertainty principle an observer effect: it's an intrinsic property of a quantum system.

\end{document}

### 2018-04-25 Articles/AMathematicalFrameworkForInformalLogic/article.tex
Notes for new articles

\documentclass[letterpaper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}

% Temporary to remove "submitted to"
\makeatletter
\def\ps@pprintTitle{%
	\let\@oddhead\@empty
	\let\@evenhead\@empty
	\def\@oddfoot{}%
	\let\@evenfoot\@oddfoot}
\makeatother

\usepackage{lineno,hyperref}

\usepackage{amsmath, amsthm, amsfonts}
\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}

\newenvironment{rationale}{\emph{Rationale}.}{\qed}
\newenvironment{justification}{\emph{Justification}.}{\qed}
\renewenvironment{proof}{\emph{Proof}.}{\qed}

\theoremstyle{plain}% default 
\newtheorem{thrm}{Theorem}[section] 
\newtheorem{prop}[thrm]{Proposition} 
\newtheorem{coro}[thrm]{Corollary}

\theoremstyle{definition}
\newtheorem{defn}[thrm]{Definition}
\newtheorem{axiom}[thrm]{Axiom}
\newtheorem*{principle*}{Principle of scientific objectivity}
\theoremstyle{remark}
\newtheorem*{rem}{Remark}

% Remove line spaces between items of enumerate and itemize
\usepackage{enumitem}
\setlist{noitemsep}

% LOGIC symbols
% -------------

% Allows to create negation symbols
\usepackage{MnSymbol}

\DeclareMathOperator{\truth}{truth}
\DeclareMathOperator{\possFn}{poss}
\DeclareMathOperator{\result}{result}
\DeclareMathOperator{\idFn}{id}

\def\TRUE{\textsc{true}}
\def\FALSE{\textsc{false}}

\def\SUCCESS{\textsc{success}}
\def\FAILURE{\textsc{failure}}
\def\UNDEF{\textsc{undefined}}

% Symbols for statements set
\def\stmtSet{\mathcal{S}}
\def\vstmtSet{\mathcal{S}_\textsf{v}}
\def\dstmtSet{\mathcal{S}_\textsf{d}}

% Symbols for tautology and contradiction
\def\tautology{\top}
\def\contradiction{\bot}

% Symbols for "compatibility" and "incompatibility"
\def\comp{\doublefrown}
\def\ncomp{\ndoublefrown}

% Symbols for "narrower" and "wider"
\def\narrower{\preccurlyeq}
\def\nnarrower{\npreccurlyeq}
\def\broader{\succcurlyeq}
\def\nbroader{\nsucccurlyeq}

% Symbol for "independent" and "correlated"
\def\indep{\upmodels}
\def\nindep{\nupmodels}

% Aliases for logical operations
\def\AND{\wedge}
\def\bigAND{\bigwedge}
\def\OR{\vee}
\def\bigOR{\bigvee}
\def\NOT{\neg}

% Formatting for statements
\newcommand{\stmt}[1][s] {\mathsf{#1}}
% Formatting for experimental tests
\newcommand{\expt}[1][e] {\mathsf{#1}}
\newcommand{\exptSet}{\mathcal{E}}
% Formatting for observations
\newcommand{\obs}[1][s] {\mathsf{#1}}
% Formatting for possibilities

% Formatting for experimental domain
\newcommand{\edomain}[1][D] {\mathcal{#1}}

% Formatting for theoretical domain
\newcommand{\tdomain}[1][D] {\bar{\mathcal{#1}}}

\newcommand{\basis}[1][B] {\mathcal{#1}} % Basis

% Formatting for experimental relationships
\newcommand{\erel}[1][r] {#1}

% Formatting for sentence statements
\newcommand{\statement}[1] {\emph{``#1''}}

\begin{document}

\title{A Mathematical Framework for Informal Logic}
\author{Gabriele Carcassi, Christine A. Aidala}

\date{\today}

\maketitle

\begin{abstract}
	In this work we present a formal framework suitable to describe the logic of informal statements. That is, the truth bearer are the abstract assertions contained in natural language while the relationships between them are still described in a formal mathematical language. This framework is guaranteed to not have antinomies, that is paradoxes that are characterized by self-inconsistencies. All paradoxes are reduced to either veridical or falsidical.
\end{abstract}

%\linenumbers

\section{Introduction}

\section{A formal logic of informal statements}

\begin{defn}
	The \textbf{Boolean domain} is the set $\mathbb{B} = \{\FALSE, \TRUE\}$ of all possible truth values.
\end{defn}

\begin{axiom}\label{ax_statement}
	A \textbf{statement} $\stmt$ is an assertion that is either true or false. Formally, a statement is an element of the set $\mathcal{S}$ of all statements upon which is defined a function $\truth: \mathcal{S} \to \mathbb{B}$ that returns the truth value for each element.
\end{axiom}

\begin{defn}
	Given a collection of statements $\{\stmt_i\}^n_{i=1}$, a \textbf{consistent truth assignment} is a collection of truth values $\{t_i\}^n_{i=1}$ such that it is logically consistent to simultaneously suppose that $\truth(\stmt_i) = t_i$ for all $1 \leq i \leq n$. That is, from those assumptions it cannot be proven that $\truth(\stmt_i) \neq t_i$ for any $1 \leq i \leq n$.  This definition generalizes to the case of infinite, possibly uncountable, collections.
\end{defn}

\begin{axiom}\label{ax_possibilities}
	The \textbf{possibilities} of a statement $\stmt$ are the possible truth values allowed by the content of the statement. Formally, on the set $\mathcal{S}$ of all statements is also defined a function $\possFn: \mathcal{S} \to \{\{\FALSE, \TRUE\},\{\FALSE\},\{\TRUE\}\}$ such that:
	\begin{itemize}
		\item $\truth(\stmt) \in \possFn(\stmt)$ for all $\stmt \in \mathcal{S}$. This remains valid in every consistent truth assignment.
		\item for any collection of statements $\{\stmt_i\}^n_{i=1}$, for any $1 \leq j \leq n$ and for any $t \in \possFn(\stmt_j)$ there exists a consistent truth assignment $\{t_i\}^n_{i=1}$ such that $t_j = t$. This generalizes to the case of infinite, possibly uncountable, indexed families.
	\end{itemize}
\end{axiom}

\begin{defn}
	A \textbf{tautology} $\tautology$ is a statement that must be true simply because of its content. That is, $\possFn(\stmt) = \{\TRUE\}$.
\end{defn}

\begin{defn}
	A \textbf{contradiction} $\contradiction$ is a statement that must be false simply because of its content. That is, $\possFn(\stmt) = \{\FALSE\}$.
\end{defn}

\begin{axiom}\label{ax_functions_of_statement}
	We can always construct a statement whose truth value arbitrarily depends on an arbitrary set of statements. Formally: given an arbitrary truth function $f_{\mathbb{B}} : \mathbb{B}^n \to \mathbb{B}$ there exists a function $f : \mathcal{S}^n \to \mathcal{S}$ such that
	$$\truth(f(\stmt_1, ..., \stmt_n)) = f_{\mathbb{B}}(\truth(\stmt_1), ..., \truth(\stmt_n))$$
	and the same relationship remains valid in every consistent truth assignment. This also holds in the case of infinite, possibly uncountable, arguments.
\end{axiom}

\begin{defn}
	Two statements $\stmt_1$ and $\stmt_2$ are \textbf{equivalent} $\stmt_1 \equiv \stmt_2$ if they must be equally true or false simply because of their content. Formally, $\stmt_1 \equiv \stmt_2$ if and only if $(\stmt_1 \AND \stmt_2) \OR (\NOT\stmt_1 \AND \NOT\stmt_2)$ is a tautology.
\end{defn}

\begin{prop}
	Statement equivalence satisfies the following properties:
	\begin{itemize}
		\item reflexivity: $\stmt \equiv \stmt$
		\item symmetry: if $\stmt_1 \equiv \stmt_2$ then $\stmt_2 \equiv \stmt_1$
		\item transitivity: if $\stmt_1 \equiv \stmt_2$ and $\stmt_2 \equiv \stmt_3$ then $\stmt_1 \equiv \stmt_3$
	\end{itemize}
	and is therefore an \textbf{equivalence relationship}.
\end{prop}

\begin{prop}\label{boolean_properties}
	The set of all statements $\mathcal{S}$ satisfies the following properties:
	\begin{itemize}
		\item associativity: $a \OR (b \OR c) \equiv (a \OR b) \OR c$, $a \AND (b \AND c) \equiv (a \AND b) \AND c$
		\item commutativity: $a \OR b \equiv b \OR a$, $a \AND b \equiv b \AND a$
		\item absorption: $a \OR (a \AND b) \equiv a$, $a \AND (a \OR b) \equiv a$
		\item identity: $a \OR \contradiction \equiv a
		$, $a \AND \tautology \equiv a$
		\item distributivity: $a \OR (b \AND c) \equiv (a \OR b) \AND (a \OR c)$, $a \AND (b \OR c) \equiv (a \AND b) \OR (a \AND c)$
		\item complements: $a \OR \NOT a \equiv \tautology$, $a \AND \NOT a \equiv \contradiction$
		\item De Morgan: $\NOT a \OR \NOT b \equiv \NOT (a \AND b)$, $\NOT a \AND \NOT b \equiv \NOT (a \OR b)$
	\end{itemize}
	This, by definition, means $\mathcal{S}$ is a \textbf{Boolean algebra}.
\end{prop}

\begin{defn}
	Given two statement $\stmt_1$ and $\stmt_2$, we say that:
	\begin{itemize}
		\item $\stmt_1$ \textbf{is narrower than} $\stmt_2$ (noted $\stmt_1 \narrower \stmt_2$) if $\stmt_2$ is true whenever $\stmt_1$ is true simply because of their content. That is, $\stmt_1 \AND \NOT \stmt_2 \equiv \contradiction$.
		\item $\stmt_1$ \textbf{is broader than} $\stmt_2$ (noted $\stmt_1 \broader \stmt_2$) if $\stmt_2 \narrower \stmt_1$.
		\item $\stmt_1$ \textbf{is compatible to} $\stmt_2$ (noted $\stmt_1 \comp \stmt_2$) if their content allows them to be true at the same time. That is, $\stmt_1 \AND \stmt_2 \nequiv \contradiction$.

	\end{itemize}
	The negation of these properties will be noted by $\nnarrower$, $\nbroader$ , $\ncomp$ respectively.
\end{defn}
\begin{defn}
	The elements of a set of statements $S \subseteq \mathcal{S}$ are said to be \textbf{independent} (noted $\stmt_1 \indep \stmt_2$ for a set of two) if their content is such that any combination of their possibilities is allowed. That is, $\possFn(f(S)) = f(\bigtimes\limits_{\stmt \in S} \possFn(\stmt))$ for any truth function $f : \mathbb{B}^{|S|} \to \mathbb{B}$. The negation of independence, will be noted by $\nindep$.
\end{defn}

\begin{prop}
	Statement narrowness satisfies the following properties:
	\begin{itemize}
		\item reflexivity: $s \narrower s$
		\item antisymmetry: if $s_1 \narrower s_2$ and  $s_2 \narrower s_1$ then $s_1 \equiv s_2$
		\item transitivity: if $s_1 \narrower s_2$ and $s_2 \narrower s_3$ then $s_1 \narrower s_3$
	\end{itemize}
	and is therefore a \textbf{partial order}.
\end{prop}
statement independence.

\section{Paradoxes}

\subsection{Self reference}

$\stmt[a]=$\statement{This sentence is true}

$\stmt[a]=$``$\truth(\stmt[a]) = \TRUE$''

\begin{table}[h]
	\centering
	\begin{tabular}{c|c}

		$\stmt[a]$ & $\truth(\stmt[a]) = \TRUE$ \\
		\hline
		1 & 1 \\
		\hline
		0 & 0 \\ 
	\end{tabular}
\end{table}

$\stmt[a]=$``$\truth(\stmt[a]) = \TRUE$'' in all cases. No paradox.

$\stmt[b]=$\statement{This sentence is false}

$\stmt[b]=$``$\truth(\stmt[b]) = \FALSE$''

\begin{table}[h]
	\centering
	\begin{tabular}{c|c}

		$\stmt[b]$ & $\truth(\stmt[b]) = \FALSE$ \\
		\hline
		0 & 1 \\
		\hline
		1 & 0 \\ 
	\end{tabular}
\end{table}

In no cases $\stmt[b]=$``$\truth(\stmt[b]) = \FALSE$''. $\possFn(\stmt[b]) = \{\}$ which cannot be. $\stmt[b]$ does not satisfy the axioms: it is not a statement. Falsidical paradox.

\subsection{Curry's paradox}

$\stmt[a]=$ \statement{If this sentence is true then I will win the lottery.}

$\stmt[b]=$\statement{I will win the lottery}

Two possible ways. Material implication $\stmt[a]=\NOT \stmt[a] \OR \stmt[b]$

\begin{table}[h]
	\centering
	\begin{tabular}{c|c|c}

		$\stmt[a]$ & $\stmt[b]$ & $\NOT \stmt[a] \OR \stmt[b]$ \\
		\hline
		0 & 0 & 1 \\
		\hline
		0 & 1 & 1 \\
		\hline
		1 & 0 & 0 \\ 
		\hline
		1 & 1 & 1 \\ 
	\end{tabular}
\end{table}

Since we must have $\stmt[a]=\NOT \stmt[a] \OR \stmt[b]$, the last line is the only one possible. Therefore $\possFn(\stmt[a])=\possFn(\stmt[b])=\{\TRUE\}$ and $\stmt[a]\equiv\stmt[b]\equiv\tautology$. If $\stmt[b]$ is not a tautology, then we have a falsidical paradox.

Other way, logical implication. $\stmt[a]=\stmt[a] \narrower \stmt[b]$.

Suppose $\stmt[a] \narrower \stmt[b]$, then the third line is ruled out because $\NOT \stmt[a] \OR \stmt[b]$ is a tautology and the first two lines are also ruled out because $\stmt[a]=\stmt[a] \narrower \stmt[b]$. Therefore, like before, $\stmt[a]\equiv\stmt[b]\equiv\tautology$. Suppose $\stmt[a] \nnarrower \stmt[b]$. Then the third line must be possible, but that can't happen because $\stmt[a]$ is false.

Even with logical implication $\stmt[a]\equiv\stmt[b]\equiv\tautology$ and if $\stmt[b]$ is not a tautology, then we have a falsidical paradox.

$\stmt[a]=$\statement{This sentence is true}

$\stmt[a] \equiv \tautology$ as seen before

$\possFn(\stmt[b]) = \{\TRUE, \FALSE\}$

$\stmt = \NOT \stmt[a] \OR \stmt[b]$

\bibliography{bibliography}

\end{document}

### 2018-04-25 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Notes for new articles

\documentclass[letterpaper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}

% Temporary to remove "submitted to"
\makeatletter
\def\ps@pprintTitle{%
	\let\@oddhead\@empty
	\let\@evenhead\@empty
	\def\@oddfoot{}%
	\let\@evenfoot\@oddfoot}
\makeatother

\usepackage{lineno,hyperref}

\usepackage{amsmath, amsthm, amsfonts}
\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}

\newenvironment{rationale}{\emph{Rationale}.}{\qed}
\newenvironment{justification}{\emph{Justification}.}{\qed}
\renewenvironment{proof}{\emph{Proof}.}{\qed}

\theoremstyle{plain}% default 
\newtheorem{thrm}{Theorem}[section] 
\newtheorem{prop}[thrm]{Proposition} 
\newtheorem{coro}[thrm]{Corollary}

\theoremstyle{definition}
\newtheorem{defn}[thrm]{Definition}
\newtheorem{axiom}[thrm]{Axiom}
\newtheorem*{principle*}{Principle of scientific objectivity}
\theoremstyle{remark}
\newtheorem*{rem}{Remark}

% Remove line spaces between items of enumerate and itemize
\usepackage{enumitem}
\setlist{noitemsep}

% LOGIC symbols
% -------------

% Allows to create negation symbols
\usepackage{MnSymbol}

\DeclareMathOperator{\truth}{truth}
\DeclareMathOperator{\possFn}{poss}
\DeclareMathOperator{\result}{result}
\DeclareMathOperator{\idFn}{id}

\def\TRUE{\textsc{true}}
\def\FALSE{\textsc{false}}

\def\SUCCESS{\textsc{success}}
\def\FAILURE{\textsc{failure}}
\def\UNDEF{\textsc{undefined}}

% Symbols for statements set
\def\stmtSet{\mathcal{S}}
\def\vstmtSet{\mathcal{S}_\textsf{v}}
\def\dstmtSet{\mathcal{S}_\textsf{d}}

% Symbols for tautology and contradiction
\def\tautology{\top}
\def\contradiction{\bot}

% Symbols for "compatibility" and "incompatibility"
\def\comp{\doublefrown}
\def\ncomp{\ndoublefrown}

% Symbols for "narrower" and "wider"
\def\narrower{\preccurlyeq}
\def\nnarrower{\npreccurlyeq}
\def\broader{\succcurlyeq}
\def\nbroader{\nsucccurlyeq}

% Symbol for "independent" and "correlated"
\def\indep{\upmodels}
\def\nindep{\nupmodels}

% Aliases for logical operations
\def\AND{\wedge}
\def\bigAND{\bigwedge}
\def\OR{\vee}
\def\bigOR{\bigvee}
\def\NOT{\neg}

% Formatting for statements
\newcommand{\stmt}[1][s] {\mathsf{#1}}
% Formatting for experimental tests
\newcommand{\expt}[1][e] {\mathsf{#1}}
\newcommand{\exptSet}{\mathcal{E}}
% Formatting for observations
\newcommand{\obs}[1][s] {\mathsf{#1}}
% Formatting for possibilities

% Formatting for experimental domain
\newcommand{\edomain}[1][D] {\mathcal{#1}}

% Formatting for theoretical domain
\newcommand{\tdomain}[1][D] {\bar{\mathcal{#1}}}

\newcommand{\basis}[1][B] {\mathcal{#1}} % Basis

% Formatting for experimental relationships
\newcommand{\erel}[1][r] {#1}

% Formatting for sentence statements
\newcommand{\statement}[1] {\emph{``#1''}}

\begin{document}

\title{Towards a general mathematical theory of experimental science}
\author{Gabriele Carcassi, Christine A. Aidala}

\date{\today}

\maketitle

\begin{abstract}
	In this work we lay the ground work for a mathematical framework that necessarily applies to all sciences. The starting point will be the logic of verifiable statements: those assertion that can be shown to be true with an experimental test. We study the algebra of such objects and show how it is closed only on finite conjunction and countable disjunction. We collect verifiable statements into experimental domains, which represent the list of all possible testable answers to a particular scientific question. From these we construct theoretical domains, which represent all possible predictions which are not necessarily verifiable themselves. We then define the possibilities of an experimental domain as those theoretical statements that, if true, predict the truthfulness of all verifiable statements in the domain. We then show that experimental domains provide a natural Kolmogorov and second countable topology over their possibilities, while the theoretical domains provide a natural $\sigma$-algebra. This gives a clear physical meaning to those mathematical structure and provide a strong justification as to why these can be used in science. It is our hope and belief that such approach can be extended to many other areas of fundamental physics.
\end{abstract}

%\linenumbers

\section{Introduction}

During the course of the last century a few fields of science and engineering have developed an associated general theory. For example, in computer science theory one starts by defining what a computational device is and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input. In communications one defines what information is and shows that no encoding can outperform the Huffman coding.\footnote{Assuming a symbol-by-symbol coding with a known input probability distribution.} In control theory one defines what a system is and then looks for strategies for optimal control. These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices. That is: they apply because of the mere definition of the subject matter.

We believe a similar approach can be developed for experimental science in general and physics in particular. That is: by focusing on what are the constraints mandated by the mere idea of studying a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories. We will see that any set of objects that are physically distinguishable must have cardinality no greater than continuum. We will see that defining what can be experimentally verified about them will always induce a topology that is Kolmogorov and second countable. These results are general because they apply independently of what is being studied.

One should not confuse this approach with the idea of a grand-unified theory or theory of everything. In that program, one looks for the theory of fundamental constituents from which all other theories can be derived. Our program is altogether different. Our subject matter are the scientific models themselves. Our aim is to understand what aspects of a scientific theory are already constrained by the mere fact that it is a scientific theory or by the assumptions one bakes into them.

What we present here is our attempt to create such a general mathematical theory of experimental science. It is the product of a decade spent in reverse engineering the laws of particle mechanics (cite paper) and our subsequent work on axiomatizing the starting points (cite other paper?). Our overall program is large and ambitious and a journal article cannot treat it comprehensively. We will simply give an overview of the starting points and the first few results: these parts are already well consolidated and provide a good template to understand our approach. We will only provide a sketch for the proofs, which can be found in our open access book (cite book) together with more details, discussions and insights.

\section{The principle of scientific objectivity}

The first thing we need to do is to characterize what science is and what its scope is. We do so by introduction the following:

\begin{principle*}
Science is universal, non-contradictory and evidence based.
\end{principle*}

This means that science can only study assertions that have a well defined truth value that can be verified experimentally. The issue at hand is to formally capture this informal intuition. We start with the following common definition.

\begin{defn}
	The \textbf{Boolean domain} is the set $\mathbb{B} = \{\FALSE, \TRUE\}$ of all possible truth values.
\end{defn}

Next we need to define our truth bearer. In mathematical logic, this is typically a well formed formula. This cannot work for us: what we are interested in is the meaning of the assertion and not how it is expressed. For example, \statement{this animal is a dog} and \statement{questo animale e' un cane} are the same statement express in different languages. As science is universal, it should not matter the language, units or reference system used to make an assertion. Therefore we need to create our own variation of classical logic. We start by defining our truth bearer as:

\begin{axiom}\label{ax_statement}
	A \textbf{statement} $\stmt$ is an assertion that is either true or false. Formally, a statement is an element of the set $\mathcal{S}$ of all statements upon which is defined a function $\truth: \mathcal{S} \to \mathbb{B}$ that returns the truth value for each element.
\end{axiom}

Note how the first part of the definition captures the informal meaning of what we are describing, while the second part captures the part that is formalized. This pattern will be present in most of our definition. So, in science, the statement is an assertion while for the math is just an element in some set.

The truth value is something that is generally established experimentally. But our model may constrain certain statements or statement combination to be ruled out. For example, \statement{this animal is a dog} and \statement{this animal is a cat} can both be either true or false, but the statement \statement{this animal is a cat and a dog} can never be true. So we need a way to keep track of these relationships so that we can never have contradictions or paradoxes.

\begin{defn}
	Given a collection of statements $\{\stmt_i\}^n_{i=1}$, a \textbf{consistent truth assignment} is a collection of truth values $\{t_i\}^n_{i=1}$ such that it is logically consistent to simultaneously suppose that $\truth(\stmt_i) = t_i$ for all $1 \leq i \leq n$. That is, from those assumptions it cannot be proven that $\truth(\stmt_i) \neq t_i$ for any $1 \leq i \leq n$.  This definition generalizes to the case of infinite, possibly uncountable, collections.
\end{defn}

Note that the truth assignments are happening in our head: they are not actual physical entities. There is only one truth value, the one found experimentally. Logical consistency is defined on the truth assignments and not on the truth value. Therefore we need a way to track whether the meaning of each statement allows it to be true or not.

\begin{axiom}\label{ax_possibilities}
	The \textbf{possibilities} of a statement $\stmt$ are the possible truth values allowed by the content of the statement. Formally, on the set $\mathcal{S}$ of all statements is also defined a function $\possFn: \mathcal{S} \to \{\{\FALSE, \TRUE\},\{\FALSE\},\{\TRUE\}\}$ such that:
	\begin{itemize}
		\item $\truth(\stmt) \in \possFn(\stmt)$ for all $\stmt \in \mathcal{S}$. This remains valid in every consistent truth assignment.
		\item for any collection of statements $\{\stmt_i\}^n_{i=1}$, for any $1 \leq j \leq n$ and for any $t \in \possFn(\stmt_j)$ there exists a consistent truth assignment $\{t_i\}^n_{i=1}$ such that $t_j = t$. This generalizes to the case of infinite, possibly uncountable, indexed families.
	\end{itemize}
\end{axiom}

With this axiom, we can distinguish between statements that can never be true and those that just happen to be true.

\begin{defn}
	A \textbf{tautology} $\tautology$ is a statement that must be true simply because of its content. That is, $\possFn(\stmt) = \{\TRUE\}$.
\end{defn}

\begin{defn}
	A \textbf{contradiction} $\contradiction$ is a statement that must be false simply because of its content. That is, $\possFn(\stmt) = \{\FALSE\}$.
\end{defn}

We also need to express relationships between the truth of different statements. For example, if we assign true to \statement{this animal is a dog} then we cannot assign true to \statement{this animal is not a dog}. Therefore we introduce the following:

\begin{axiom}\label{ax_functions_of_statement}
	We can always construct a statement whose truth value arbitrarily depends on an arbitrary set of statements. Formally: given an arbitrary truth function $f_{\mathbb{B}} : \mathbb{B}^n \to \mathbb{B}$ there exists a function $f : \mathcal{S}^n \to \mathcal{S}$ such that
	$$\truth(f(\stmt_1, ..., \stmt_n)) = f_{\mathbb{B}}(\truth(\stmt_1), ..., \truth(\stmt_n))$$
	and the same relationship remains valid in every consistent truth assignment. This also holds in the case of infinite, possibly uncountable, arguments.
\end{axiom}

We will use the standard symbols $\NOT$, $\AND$ $\OR$ to indicate the negation, conjunction (logical AND) and disjunction (logical OR). With these three axioms, we can rederive all of of the rules of classical logic. First we define our equivalence.

\begin{defn}
	Two statements $\stmt_1$ and $\stmt_2$ are \textbf{equivalent} $\stmt_1 \equiv \stmt_2$ if they must be equally true or false simply because of their content. Formally, $\stmt_1 \equiv \stmt_2$ if and only if $(\stmt_1 \AND \stmt_2) \OR (\NOT\stmt_1 \AND \NOT\stmt_2)$ is a tautology.
\end{defn}

From this definition, one can prove that two statements are equivalent if and only if they have the same truth in all consistent truth assignments. From that one can prove the following two propositions.

\begin{prop}
	Statement equivalence satisfies the following properties:
	\begin{itemize}
		\item reflexivity: $\stmt \equiv \stmt$
		\item symmetry: if $\stmt_1 \equiv \stmt_2$ then $\stmt_2 \equiv \stmt_1$
		\item transitivity: if $\stmt_1 \equiv \stmt_2$ and $\stmt_2 \equiv \stmt_3$ then $\stmt_1 \equiv \stmt_3$
	\end{itemize}
	and is therefore an \textbf{equivalence relationship}.
\end{prop}

\begin{prop}\label{boolean_properties}
	The set of all statements $\mathcal{S}$ satisfies the following properties:
	\begin{itemize}
		\item associativity: $a \OR (b \OR c) \equiv (a \OR b) \OR c$, $a \AND (b \AND c) \equiv (a \AND b) \AND c$
		\item commutativity: $a \OR b \equiv b \OR a$, $a \AND b \equiv b \AND a$
		\item absorption: $a \OR (a \AND b) \equiv a$, $a \AND (a \OR b) \equiv a$
		\item identity: $a \OR \contradiction \equiv a
		$, $a \AND \tautology \equiv a$
		\item distributivity: $a \OR (b \AND c) \equiv (a \OR b) \AND (a \OR c)$, $a \AND (b \OR c) \equiv (a \AND b) \OR (a \AND c)$
		\item complements: $a \OR \NOT a \equiv \tautology$, $a \AND \NOT a \equiv \contradiction$
		\item De Morgan: $\NOT a \OR \NOT b \equiv \NOT (a \AND b)$, $\NOT a \AND \NOT b \equiv \NOT (a \OR b)$
	\end{itemize}
	This, by definition, means $\mathcal{S}$ is a \textbf{Boolean algebra}.
\end{prop}

One can also prove that the Boolean algebra is complete, and therefore it generalizes to the infinite case.

This system not only allows to use formal classical logic while keeping the statements informal, but it also equipped to capture causal relationships.\footnote{Note that the axioms prevents any form of paradox, simply because the possibilities of a statement are never empty and a consistent truth assignment must exist. Therefore situations like \statement{this statement is not true} are ruled out simply because they cannot satisfy the axioms.} For example, $\stmt_1=$\statement{the thermometer indicator is between 24 C and 25 C} and $\stmt_s=$\statement{the temperature is between 24 C and 25 C} are equivalent because both $\stmt_1 \AND \NOT \stmt_2$ and $\NOT \stmt_1 \AND \stmt_2$ are contradictions (on the assumption that our thermometer is actually working). We can also define other semantic relationships.

\begin{defn}
	Given two statement $\stmt_1$ and $\stmt_2$, we say that:
	\begin{itemize}
		\item $\stmt_1$ \textbf{is narrower than} $\stmt_2$ (noted $\stmt_1 \narrower \stmt_2$) if $\stmt_2$ is true whenever $\stmt_1$ is true simply because of their content. That is, $\stmt_1 \AND \NOT \stmt_2 \equiv \contradiction$.
		\item $\stmt_1$ \textbf{is broader than} $\stmt_2$ (noted $\stmt_1 \broader \stmt_2$) if $\stmt_2 \narrower \stmt_1$.
		\item $\stmt_1$ \textbf{is compatible to} $\stmt_2$ (noted $\stmt_1 \comp \stmt_2$) if their content allows them to be true at the same time. That is, $\stmt_1 \AND \stmt_2 \nequiv \contradiction$.

	\end{itemize}
	The negation of these properties will be noted by $\nnarrower$, $\nbroader$ , $\ncomp$ respectively.
\end{defn}
\begin{defn}
	The elements of a set of statements $S \subseteq \mathcal{S}$ are said to be \textbf{independent} (noted $\stmt_1 \indep \stmt_2$ for a set of two) if their content is such that any combination of their possibilities is allowed. That is, $\possFn(f(S)) = f(\bigtimes\limits_{\stmt \in S} \possFn(\stmt))$ for any truth function $f : \mathbb{B}^{|S|} \to \mathbb{B}$. The negation of independence, will be noted by $\nindep$.
\end{defn}

For example, given the following statements:
\begin{description}
	\item $\stmt_1=$\statement{that animal is a cat}
	\item $\stmt_2=$\statement{that animal is a mammal}
	\item $\stmt_3=$\statement{that animal is a dog}
	\item $\stmt_4=$\statement{that animal is black}
\end{description}
we have the $\stmt_1$ $\narrower$ $\stmt_2$ (i.e. it is more specific), $\stmt_1$ $\ncomp$ $\stmt_3$ (i.e. they can't be true at the same time) and $\stmt_1$ $\indep$ $\stmt_4$ (i.e. the truth of one tells us nothing about the truth of the other).

Another interesting result is that statement narrowness imposes a partial order on the set of all statements.

\begin{prop}
	Statement narrowness satisfies the following properties:
	\begin{itemize}
		\item reflexivity: $s \narrower s$
		\item antisymmetry: if $s_1 \narrower s_2$ and  $s_2 \narrower s_1$ then $s_1 \equiv s_2$
		\item transitivity: if $s_1 \narrower s_2$ and $s_2 \narrower s_3$ then $s_1 \narrower s_3$
	\end{itemize}
	and is therefore a \textbf{partial order}.
\end{prop}

As every element in out theory will be constructed upon statements, these operations are important as they will characterize all those constructions. For example, if we quantify the precision of a set of statements, it will need to be ordered in a way that is compatible with narrowness. If we define statistical independence, it will have to be defined in a way that is compatible with statement independence.

\section{Verifiable statements and experimental domains}

We have the tools for dealing with assertions that are universal and non-contradictory, now we have to develop the tools for evidence based.

\begin{axiom}\label{ax_verifiable_statements}
	A \textbf{verifiable statement} is a statement that can be shown to be true experimentally. Formally, a statement $\stmt$ is verifiable if it is part of the subset $\stmt \in \vstmtSet \subset \stmtSet$ of all verifiable statements.
\end{axiom}

Physically, this means that we have a repeatable procedure that anybody can execute and that always gives the same result. If the statement is true, then this experimental test must terminate successfully. Note that, in general, the test may not terminate if the statement is false. Consider the following:
\begin{enumerate}
	\item find a swan
	\item if it's black terminate successfully
	\item go to step 1
\end{enumerate}
This will terminate if a black swan is found, but it will not terminate if no black swans exist. Because of non-termination, failure to verify is not verification of the negation. That is: the negation of a verifiable statement is not necessarily a verifiable statement.

Given two statements, though, we can verify their conjunction simply by verifying both: if they are both true, both their tests will terminate and verify the conjunction. But we cannot extend this to an infinite number of statements as we would never terminate. We can also verifying the disjunction of two statement: once one test terminates we are done. And because we only need one test to terminate, we can generalize to a countable number of verifiable statements by following this procedure:
\begin{enumerate}
	\item initialize $n$ to 1
	\item for each $i=1..n$
	\begin{enumerate}
		\item run the test for $\stmt_i$ for $n$ seconds
		\item if it terminates successfully then terminate successfully
	\end{enumerate}
	\item increment $n$ and go to step 2
\end{enumerate}
This procedure will run all tests for an arbitrary amount of time therefore, if one statement is true, it will make the test terminate successfully, even if all other tests would never terminate. In light of this, we can set the following axioms.

\begin{axiom}\label{ax_verifiable_AND}
	The conjunction of a finite collection of verifiable statements is a verifiable statement. Formally, let $\{\stmt_i\}_{i=1}^{n} \subseteq \vstmtSet$ be a finite collection of verifiable statements. Then the conjunction $\bigAND\limits_{i=1}^{n} \stmt_i \in \vstmtSet$ is a verifiable statement.
\end{axiom}
	\begin{axiom}\label{ax_verifiable_OR}
	The disjunction of a countable collection of verifiable statements is a verifiable statement. Formally, let $\{\stmt_i\}_{i=1}^{\infty} \subseteq \vstmtSet$ be a countable collection of verifiable statements. Then the disjunction $\bigOR\limits_{i=1}^{\infty} \stmt_i \in \vstmtSet$ is a verifiable statement.
\end{axiom}

We could also define decidable statements as those for which falsehood can also be tested experimentally. This table compares the different algebras.

\begin{table}[h]
	\centering
	\begin{tabular}{p{0.14\textwidth} p{0.08\textwidth} p{0.13\textwidth} p{0.22\textwidth} p{0.23\textwidth}}
		Operator & Gate & Statement & Verifiable Statement & Decidable Statement  \\ 
		\hline 
		Negation & NOT & allowed & disallowed & allowed \\ 
		Conjunction & AND & arbitrary  & finite & finite \\ 
		Disjunction & OR & arbitrary  & countable & finite \\ 
	\end{tabular}
	\caption{Comparing algebras of statements.}
\end{table}

Now that we can verify statements one by one, we need to define what it means to verify a group of them. First, we note that we may not need to run all tests for all statements. For example, if we verified $\stmt_1$ then there is no need to verify any of its disjunctions like $\stmt_1 \OR \stmt_2$.

\begin{defn}
	Given a set $\edomain$ of verifiable statements, $\basis \subseteq \edomain$ is a \textbf{basis} if the truth values of $\basis$ are enough to deduce the truth values of the set. Formally, all elements of $\edomain$ can be generated from $\basis$ using finite conjunction and countable disjunction.
\end{defn}

It is the size of the basis, then, that matters. We'll consider the case when we have an indefinitely long time at our disposal. We do that because it's the upper limit, because it handles limits and because we haven't given constraints for how long or short a successful test can be (i.e. if ran a test for three days it may still successfully complete in three days and one minute). As we can keep going, we can be given a countable set of verifiable statement and at some point, if any statement is true, we are going to verify it experimentally.

\begin{defn}
	An \textbf{experimental domain} $\edomain$ represents all the experimental evidence that can be acquired about a scientific subject in an indefinite amount of time. Formally, it is a set of statements, closed under finite conjunction and countable disjunction, that includes precisely the tautology, the contradiction, and a set of verifiable statements that can be generated from a countable basis.
\end{defn}

These axioms and definitions formally characterize what we mean by evidence based. A scientific theory will be fully defined by a countable set of verifiable statements, the basis for an experimental domain.

\section{Theoretical domain and possibilities}

While verifiable statements define what can be scientifically studied, not all predictions are themselves verifiable. Consider the two statements \statement{there exists extra-terrestrial life} and \statement{there is no extra-terrestrial life}. We can verify the first if we happen to find signs of life somewhere, but experimentally verifying the second is practically impossible. Yet, the second is still meaningful as a prediction: it predicts that the test for the first will never terminate. That is: while negations are not verifiable we can still logically talk about them when, for example, constructing truth assignments.

\begin{defn}
	The \textbf{theoretical domain} $\tdomain$ of an experimental domain $\edomain$ is the set of statements that we can use to state predictions, which is constructed from $\edomain$ by allowing negation. We call \textbf{theoretical statement} a statement that is part of a theoretical domain. More formally, $\tdomain$ is the set of all statements generated from $\edomain$ using negation, finite conjunction and countable disjunction.
\end{defn}

Note that the new statements provide no new information. In fact, all statements in the theoretical domain $\tdomain$ can be generated by negation, countable conjunction and countable disjunction from a basis $\basis$ of $\edomain$. As they provide all possible predictions, we now found the ones that completely specify the truth value for all theoretical statements. For example, once we know that \statement{this animal is a cat} we know that \statement{this animal has whiskers}, that \statement{this animal has no feathers}and so on.

\begin{defn}
	A \textbf{possibility} for an experimental domain $\edomain$ is a statement $x \in \tdomain$ that, when true, determines the truth value for all statements in the theoretical domain. Formally, $x \nequiv \contradiction$ and for each $\mathsf{s} \in \tdomain$, either $x \narrower \mathsf{s}$ or $x \ncomp \mathsf{s}$. The \textbf{possibilities} $X$ for $\edomain$ are the collection of all possibilities.
\end{defn}

The possibilities are all the different cases that can be distinguished experimentally with the verifiable statements of the domain. We conclude this section with the following general result.

\begin{thrm}
	The possibilities $X$ for an experimental domain $\edomain$ have at most the cardinality of the continuum.
\end{thrm}

The proof is simply noting that each possibility can be labeled by the truth value of the countable basis of the experimental domain. We cannot have more possibilities than countable sequences of true/false and the set of all binary countable sequences has the cardinality of continuum.

This means that we are never going to be able to experimentally distinguish between elements of greater cardinality. The set of discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$, for example, has a greater cardinality and therefore it will never describe anything that is physically relevant. All the issues with large cardinals are not something science will ever be interested in. It does not matter what system we are describing, what experimental techniques we are using or how clever we are.

\section{Topologies and sigma-algebras}

Now that we have introduced the basic mathematical structures for experimental science, we show that their deep connection to other well established mathematical structures.

We first note that each verifiable statement can be written as the disjunction of a set of possibilities.

\begin{defn}
	Let $\edomain$ be an experimental domain and $X$ its possibilities. We define the map $U : \edomain \rightarrow 2^X$ that for each statement $\obs \in \edomain$ returns the set of possibilities compatible with it. That is: $U(\obs)\equiv\{ x \in X \, | \, x \comp \obs\}$. We call $U(\obs)$ the \textbf{verifiable set} of possibilities associated with $\obs$.
\end{defn}

\begin{prop}
	A statement $\obs \in \edomain$ is the disjunction of the possibilities in its verifiable set $U(\obs)$. That is, $\obs=\bigOR\limits_{x \in U(\obs)} x$.
\end{prop}

The proof is a trivial application of the disjunction normal form of Boolean algebra. In fact, each possibility can be written as a minterm of a basis (i.e. a conjunction where each basis element appears only once either negated or not). Any verifiable statement can be expressed in terms of the basis, and it is a result of Boolean algebra that each logical expression can be formulated as a disjunction of minterms (i.e. an OR of ANDs). Intuitively, it is the disjunction of all cases in which the statement is true.

Since every statement is a set of possibilities, we can re-express the statement relationships in terms of sets according to this table:

\begin{table}[h]
	\centering
	\begin{tabular}{p{0.075\textwidth} p{0.275\textwidth} p{0.2\textwidth} p{0.3\textwidth}}
		& Statement relationship & & Set relationship  \\ 
		\hline 
		$\stmt_1 \AND \stmt_2$ & (Conjunction) & $U(\stmt_1) \cap U(\stmt_2)$ & (Intersection) \\ 
		$\stmt_1 \OR \stmt_2$ & (Disjunction) & $U(\stmt_1) \cup U(\stmt_2)$ & (Union) \\ 
		$\NOT \stmt$ & (Negation) & $U(\stmt)^C$ & (Complement) \\ 
		$\stmt_1 \equiv \stmt_2$ & (Equivalence) & $U(\stmt_1) = U(\stmt_2)$ & (Equality) \\ 
		$\stmt_1 \narrower \stmt_2$ & (Narrower than) & $U(\stmt_1) \subseteq U(\stmt_2)$ & (Subset) \\ 
		$\stmt_1 \broader \stmt_2$ & (Broader than) & $U(\stmt_1) \supseteq U(\stmt_2)$ & (Superset) \\ 
		$\stmt_1 \comp \stmt_2$ & (Compatibility) & $U(\stmt_1) \cap U(\stmt_2) \neq \emptyset$ & (Intersection not empty)
	\end{tabular} 
	\caption{Correspondence between statement operators and set operators.}
\end{table}

The closure of an experimental domain under finite conjunction and countable disjunction becomes closure under finite intersection and countable union. Since the basis is countable, countable union is equivalent to arbitrary union. In other words: the set of all verifiable sets is a topology.

\begin{thrm}
	Let $X$ be the set of possibilities for an experimental domain $\edomain$. $X$ has a natural topology given by the collection of all verifiable sets $\mathsf{T}_X=U(\edomain)$ that is Kolmogorov and second countable.
\end{thrm}

The topology is Kolmogorov (i.e. $T_0$) because given two possibilities, by their construction, there must be one element of the basis that is compatible with one but not the other. It is second countable because a basis of the experimental domain corresponds to a sub-basis of the topology.

One can also show that the natural topology is Hausdorff if and only if all possibilities are approximately verifiable. That is, there exists a sequence $\{\obs_i\}_{i=1}^{\infty} \in \edomain$ such that $x = \bigAND\limits_{i=1}^{\infty} \obs_i$ for each $x \in X$.

In the same way, the theoretical domain corresponds to a $\sigma$-algebra on the possibilities.

\begin{defn}
	Let $\tdomain$ be a theoretical domain and $X$ its possibilities. We define the map $A : \tdomain \rightarrow 2^X$ that for each theoretical statement $\stmt \in \tdomain$ returns the set of possibilities compatible with it. That is, $A(\stmt)\equiv\{ x \in X \, | \, x \comp \stmt\}$. We call $A(\stmt)$ the \textbf{theoretical set} of possibilities associated with $\stmt$
\end{defn}

\begin{thrm}
	Let $X$ be the set of possibilities for a theoretical domain $\tdomain$. $X$ has a natural $\sigma$-algebra given by the collection of all theoretical sets $\Sigma_X=A(\tdomain)$.
\end{thrm}

The proof is again a simple mapping of statement operations to set operations. It can also be shown that the natural $\sigma$-algebra of the possibilities is the Borel algebra of their natural topology.

To give an example, consider an experimental domain formed by statements like \statement{this quantity is between $a$ and $b$}. For example, \statement{the distance between the earth and the moon is between 384 and 385 thousand Km}. And suppose that the precision can be arbitrarily narrowed. We can define the following.

\begin{defn}
	An experimental domain $\edomain_X$ is said to allow \textbf{arbitrary precision} if it describes a quantity measurable with arbitrary precision. Formally, there exists a countable basis $\basis \subseteq \edomain_X$ such that:
	\begin{itemize}
		\item we can define a function $\operatorname{bounds} : \basis \to \mathbb{Q} \times \mathbb{Q}$ such that $\operatorname{bounds}(\stmt) = (a, b)$ is a tuple where $a < b$ for all $\stmt \in \basis$
		\item for every pair of rational numbers $(a, b) \in \mathbb{Q} \times \mathbb{Q}$ there exists an $\obs \in \basis$ such that $\operatorname{bounds}(\stmt) = (a, b)$
		\item if $\stmt = \stmt_1 \AND \stmt_2$, $\operatorname{bounds}(\stmt_1) = (a_1, b_1)$ and $\operatorname{bounds}(\stmt_2) = (a_2, b_2)$, then $\stmt \equiv \contradiction$ if $b_1 \leq a_2$ or $b_2 \leq a_1$ otherwise $\stmt \nequiv \contradiction$ and $\operatorname{bounds}(\stmt) = (\max(a_1, a_2), \min(b_1, b_2))$
	\end{itemize}
\end{defn}

Then it can be shown that:

\begin{thrm}[Arbitrary precision is continuity]\label{thrm_arbitrary_precision_is_continuity}
	Given an experimental domain $\edomain_X$, the set of possibilities with their natural topology is homeomorphic to the reals with the standard topology if and only if the domain allows arbitrary precision.
\end{thrm}

This can be proven by showing that each possibility is equivalent to a class of Cauchy sequences, which corresponds to a real number. Each open interval corresponds to a finite precision measurement. The possibilities are not themselves verifiable statements: we cannot measure a continuous quantity with infinite precision. But it is a theoretical statement: we can use the value to predict which finite precision measurement will be verified (the ones that contain that value).

This connection creates a strong bridge between the standard mathematical structures used in physics and their meaning. Every theorem, every proof now has a consistent and direct physical meaning. And given the foundational nature of topologies and $\sigma$-algebra, we have hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.

\section{Conclusion}

The theory presented in this paper can be quickly summarized as follows. Science deals with statements that can be shown to be true experimentally. A scientific domain, then, will be defined by a set of verifiable statements, an experimental domain. From these we construct the set of all possible predictions, the theoretical domain, and concentrate on those that fully predict all our measurements, the possibilities. By the very nature and properties of these objects, a topology and a $\sigma$-algebra is induced on the possibilities that retains the connection to what statements are verifiable and what statements constitute a valid prediction.

The framework can be further be expanded by studying how additional assumptions constrain the verifiable statements. We have seen, for example, how assuming a quantity can be measured with arbitrary precision leads to the real numbers with standard topology. Similarly, other assumptions will lead to other mathematical structures.

We feel that this type of work is of significance importance for the scientific community in general and for the physics community in particular if we are ever to fully understand what scientific theories are and what are they meant to be describing.

\bibliography{bibliography}

\end{document}

### 2018-05-02 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Review of the math phys article

@@
 For example, in computer science theory one starts by defining what a computational device is and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input.
-In communications one defines what information is and shows that no encoding can outperform the Huffman coding.\footnote{Assuming a symbol-by-symbol coding with a known input probability distribution.} In control theory one defines what a system is and then looks for strategies for optimal control.
+In communications one defines what information is and shows that no encoding can outperform the Huffman coding.\footnote{Assuming a symbol-by-symbol coding with a known input probability distribution.} In control theory one defines what a system is and then looks for general strategies for control such as Kalman filtering.
 These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
-That is: they apply because of the mere definition of the subject matter.
+That is: the results apply because of the mere definition of the subject matter.
 
-We believe a similar approach can be developed for experimental science in general and physics in particular.
-That is: by focusing on what are the constraints mandated by the mere idea of studying a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories.
-We will see that any set of objects that are physically distinguishable must have cardinality no greater than continuum.
+No such general theory currently exists for physics, which is essentially composed of different theories loosely connected to each other, each with their own realm of applicability.
+However, we believe it can be created and in this work we present our basic strategy towards it.
+As one familiar to the general theories of other subjects would expect, we only need to focus on clarifying exactly what our starting points are and let the logic follow.
+That is: once we formalize exactly what it means to study a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories.
+For example, we will see that any set of objects that are physically distinguishable must have cardinality no greater than continuum.
 We will see that defining what can be experimentally verified about them will always induce a topology that is Kolmogorov and second countable.
@@
 In that program, one looks for the theory of fundamental constituents from which all other theories can be derived.
-Our program is altogether different.
+Our program is altogether different and orthogonal.
 Our subject matter are the scientific models themselves.
 Our aim is to understand what aspects of a scientific theory are already constrained by the mere fact that it is a scientific theory or by the assumptions one bakes into them.
+While this general theory may be of some utility to someone working on a grand-unified theory as it may help narrow the search, neither one will lead to the other.
 
-What we present here is our attempt to create such a general mathematical theory of experimental science.
+In this paper we briefly present the starting point for our general mathematical theory of experimental science.
 It is the product of a decade spent in reverse engineering the laws of particle mechanics (cite paper) and our subsequent work on axiomatizing the starting points (cite other paper?).
 Our overall program is large and ambitious and a journal article cannot treat it comprehensively.
-We will simply give an overview of the starting points and the first few results: these parts are already well consolidated and provide a good template to understand our approach.
+Therefore we will only present an overview of the basic concepts, which are already well consolidated and provide a good template to understand our approach.
 We will only provide a sketch for the proofs, which can be found in our open access book (cite book) together with more details, discussions and insights.
+
+The basic idea is that science deals with verifiable statements: assertion whose truth can be tested experimentally.
+We study the logic of these statements, which is different from the standard Boolean logic.
+We then group them into experimental domains, which represent a set of verifiable statements that can be tested in an indefinite amount of time.
+From these we construct theoretical domains, which represent all possible predictions, and within these we find the possibilities, those predictions that determine the outcome of all possible tests.
+We will find that possibilities have a natural topology induced by the experimental domain and a natural $\sigma$-algebra induced by the theoretical domain.
+We will find that the topology is Kolmogorov and second countable, which limits the cardinality of the set of possibilities of a domain to the one of continuum.
+Note that these structure are indeed pervasive in science, and form the foundation for differential geometry (Riemannian and symplectic), measure theory, probability theory and many other mathematical tools already used in physics and other sciences.
 
@@
 
-This means that science can only study assertions that have a well defined truth value that can be verified experimentally.
+This means that science restricts itself to the study of assertions that have a well defined truth value which can be verified experimentally.
 The issue at hand is to formally capture this informal intuition.
@@
 This cannot work for us: what we are interested in is the meaning of the assertion and not how it is expressed.
-For example, \statement{this animal is a dog} and \statement{questo animale e' un cane} are the same statement express in different languages.
+For example, \statement{this animal is a dog} and \statement{questo animale \`e un cane} are the same statement express in different languages.
 As science is universal, it should not matter the language, units or reference system used to make an assertion.
-Therefore we need to create our own variation of classical logic.
+Therefore we introduce a variation of algebraic logic suited to our needs.
 We start by defining our truth bearer as:
@@
 Note how the first part of the definition captures the informal meaning of what we are describing, while the second part captures the part that is formalized.
-This pattern will be present in most of our definition.
-So, in science, the statement is an assertion while for the math is just an element in some set.
+This pattern will be present in most of our definition and it serves to clarify both what is being formalized and how.
+Therefore, in science, the statement is an assertion while for the math is just an element in some set.
 
-The truth value is something that is generally established experimentally.
+While in math the truth value is the focus of the logic system, in science it is generally established experimentally.
 But our model may constrain certain statements or statement combination to be ruled out.
 For example, \statement{this animal is a dog} and \statement{this animal is a cat} can both be either true or false, but the statement \statement{this animal is a cat and a dog} can never be true.
-So we need a way to keep track of these relationships so that we can never have contradictions or paradoxes.
+That is: the role of our logic system is not to keep track of what is true and false, but of what can't possibly be true and can't possibly be false.
+So we need to keep track of these relationships so that we can never have inconsistencies or paradoxes.
 
@@
 That is, from those assumptions it cannot be proven that $\truth(\stmt_i) \neq t_i$ for any $1 \leq i \leq n$.
-This definition generalizes to the case of infinite, possibly uncountable, collections.
+This definition generalizes to the case of infinite, possibly uncountable, indexed families.
 \end{defn}
@@
 	The \textbf{possibilities} of a statement $\stmt$ are the possible truth values allowed by the content of the statement.
-Formally, on the set $\mathcal{S}$ of all statements is also defined a function $\possFn: \mathcal{S} \to \{\{\FALSE, \TRUE\},\{\FALSE\},\{\TRUE\}\}$ such that:
+Formally, on $\mathcal{S}$ is also defined a function $\possFn: \mathcal{S} \to \{\{\FALSE, \TRUE\},\{\FALSE\},\{\TRUE\}\}$ such that:
 	\begin{itemize}
@@
 
-We will use the standard symbols $\NOT$, $\AND$ $\OR$ to indicate the negation, conjunction (logical AND) and disjunction (logical OR).
+We will use the standard symbols $\NOT$, $\AND$ $\OR$ to indicate the negation (logical NOT), conjunction (logical AND) and disjunction (logical OR).
 With these three axioms, we can rederive all of of the rules of classical logic.
@@
 
-As every element in out theory will be constructed upon statements, these operations are important as they will characterize all those constructions.
+As every element in our general theory will be constructed upon statements, these operations are important as they will characterize all those constructions.
 For example, if we quantify the precision of a set of statements, it will need to be ordered in a way that is compatible with narrowness.
 If we define statistical independence, it will have to be defined in a way that is compatible with statement independence.
+In the same way that set theory defines concepts common on all mathematics, the general theory defines basic concepts that are common to all scientific theories.
 
@@
 
-We have the tools for dealing with assertions that are universal and non-contradictory, now we have to develop the tools for evidence based.
+We have the tools for dealing with assertions that are universal and non-contradictory, now we have to develop the tools for those that areevidence based as well.
 
@@
 \end{enumerate}
-This will terminate if a black swan is found, but it will not terminate if no black swans exist.
+This will terminate if a black swan is found, but it will not terminate if no black swans exist (i.e.
+absence of evidence is not evidence of absence).
 Because of non-termination, failure to verify is not verification of the negation.
@@
 Now that we can verify statements one by one, we need to define what it means to verify a group of them.
-First, we note that we may not need to run all tests for all statements.
-For example, if we verified $\stmt_1$ then there is no need to verify any of its disjunctions like $\stmt_1 \OR \stmt_2$.
+In principle, given a set of verifiable statements we can simply start testing them one after the other.
+However, if we are given a set of uncountable statements than, even if we have an indefinitely long time at our disposal, we will not be able to create a procedure that eventually tests all statements.
+However, we may not need to actually run all tests for all statements.
+For example, if we found that $\stmt_1$ is true then there is no need to test any of its disjunctions like $\stmt_1 \OR \stmt_2$.
 
@@
 
-It is the size of the basis, then, that matters.
-We'll consider the case when we have an indefinitely long time at our disposal.
-We do that because it's the upper limit, because it handles limits and because we haven't given constraints for how long or short a successful test can be (i.e.
-if ran a test for three days it may still successfully complete in three days and one minute).
-As we can keep going, we can be given a countable set of verifiable statement and at some point, if any statement is true, we are going to verify it experimentally.
+Therefore it is the size of the basis, then, that matters and not the size of the set.
+If the basis is countable we can keep going and, if any statement is true, it will eventually be verified experimentally.\footnote{We assume we are given an indefinitely long time that because it's the upper limit, because it handles limits and because we haven't given constraints for how long or short a successful test can be (i.e.
+if ran a test for three days it may still successfully complete in three days and one minute).}
 
@@
 
-\section{Theoretical domain and possibilities}
+\section{Theoretical domains and possibilities}
 
@@
 In fact, all statements in the theoretical domain $\tdomain$ can be generated by negation, countable conjunction and countable disjunction from a basis $\basis$ of $\edomain$.
-As they provide all possible predictions, we now found the ones that completely specify the truth value for all theoretical statements.
+As they provide all possible predictions, we focus on the ones that completely specify the truth value for all theoretical statements.
 For example, once we know that \statement{this animal is a cat} we know that \statement{this animal has whiskers}, that \statement{this animal has no feathers}and so on.
@@
 
-The possibilities are all the different cases that can be distinguished experimentally with the verifiable statements of the domain. We conclude this section with the following general result.
+The possibilities are all the different cases that can be distinguished experimentally given the verifiable statements of the domain.
+If we increase or otherwise change the set of verifiable statements (e.g.
+we learn how to test the DNA of animals) then the possibilities will change as well (e.g.
+the possible animal species is refined).
+We conclude this section with the following general result.
 
@@
 The proof is simply noting that each possibility can be labeled by the truth value of the countable basis of the experimental domain.
-We cannot have more possibilities than countable sequences of true/false and the set of all binary countable sequences has the cardinality of continuum.
+We cannot have more possibilities than sequences of true/false and the set of all binary sequences has the cardinality of continuum.
 
 This means that we are never going to be able to experimentally distinguish between elements of greater cardinality.
-The set of discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$, for example, has a greater cardinality and therefore it will never describe anything that is physically relevant.
+The set of discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$, for example, has greater cardinality and therefore it will never be associated with any physically distinguishable concept.
 All the issues with large cardinals are not something science will ever be interested in.
@@
 
-Now that we have introduced the basic mathematical structures for experimental science, we show that their deep connection to other well established mathematical structures.
+Now that we have introduced the basic mathematical structures for our general theory, we show that their deep connection to other well established mathematical structures.
 
@@
 
-One can also show that the natural topology is Hausdorff if and only if all possibilities are approximately verifiable.
-That is, there exists a sequence $\{\obs_i\}_{i=1}^{\infty} \in \edomain$ such that $x = \bigAND\limits_{i=1}^{\infty} \obs_i$ for each $x \in X$.
+One can also show that the natural topology is Hausdorff if and only if all possibilities are approximately verifiable (i.e.
+each possibility is the limit of a sequence of verifiable statements).
+The hope is that we can find physically meaningful definitions for all topological concepts.
 
@@
 
-The theory presented in this paper can be quickly summarized as follows.
+The general theory presented in this paper can be quickly summarized as follows.
 Science deals with statements that can be shown to be true experimentally.
-A scientific domain, then, will be defined by a set of verifiable statements, an experimental domain.
-From these we construct the set of all possible predictions, the theoretical domain, and concentrate on those that fully predict all our measurements, the possibilities.
+A scientific domain (i.e.
+an experimental domain) will be defined by a set of verifiable statements.
+From these we construct the set of all possible predictions (i.e.
+the theoretical domain), and concentrate on those that fully predict all our measurements, the possibilities.
 By the very nature and properties of these objects, a topology and a $\sigma$-algebra is induced on the possibilities that retains the connection to what statements are verifiable and what statements constitute a valid prediction.
@@
 Similarly, other assumptions will lead to other mathematical structures.
+This expansion is the aim of current and future work.
 
-We feel that this type of work is of significance importance for the scientific community in general and for the physics community in particular if we are ever to fully understand what scientific theories are and what are they meant to be describing.
+One benefit of the general theory is to bring to the forefront our hidden assumptions.
+We have seen that real numbers are associated with arbitrary precision.
+What happens in the case of limited precision, when we cannot assume precision can always be increased indefinitely?
+Would we still have real numbers?
+Would we be justified to use smooth manifolds or would we need an entirely different topological structure?
+
+The other benefit is that requires no interpretation.
+All mathematical definition are capturing a well defined physical concept.
+Therefore every theorem, every step in every proof will also have a well defined physical meaning.
+Since we know what we have put in, we know what comes out.
+
+We feel that this type of work is of significance importance for the scientific community in general and for the physics community in particular if we are ever to truly understand what scientific theories are and what are they meant to be describing.
 

### 2018-05-11 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Finalizing the first version

@@
 \title{Towards a general mathematical theory of experimental science}
-\author{Gabriele Carcassi, Christine A. Aidala}
+\author{Gabriele Carcassi, Christine A. Aidala \\ University of Michigan}
 
@@
 \begin{abstract}
-	In this work we lay the ground work for a mathematical framework that necessarily applies to all sciences.
-The starting point will be the logic of verifiable statements: those assertion that can be shown to be true with an experimental test.
+	In this article we lay the ground work for a general mathematical theory of experimental science.
+The starting point will be the notion of verifiable statements, those assertions that can be shown to be true with an experimental test.
 We study the algebra of such objects and show how it is closed only on finite conjunction and countable disjunction.
-We collect verifiable statements into experimental domains, which represent the list of all possible testable answers to a particular scientific question.
-From these we construct theoretical domains, which represent all possible predictions which are not necessarily verifiable themselves.
-We then define the possibilities of an experimental domain as those theoretical statements that, if true, predict the truthfulness of all verifiable statements in the domain.
-We then show that experimental domains provide a natural Kolmogorov and second countable topology over their possibilities, while the theoretical domains provide a natural $\sigma$-algebra.
-This gives a clear physical meaning to those mathematical structure and provide a strong justification as to why these can be used in science.
-It is our hope and belief that such approach can be extended to many other areas of fundamental physics.
+With simple constructions, we show that the set of possible cases distinguishable by verifiable statements is equipped with a natural Kolmogorov and second countable topology and a natural $\sigma$-algebra.
+This gives a clear physical meaning to those mathematical structures and provides a strong justification for their use in science.
+It is our hope and belief that such an approach can be extended to many areas of fundamental physics and will provide a consistent vocabulary across scientific domains.
 \end{abstract}
+
+% We form experimental domains, which represent the list of all possible testable answers to a particular scientific question.
+From these we construct theoretical domains, which represent all possible predictions (not necessarily verifiable themselves).
+We then define the possibilities of an experimental domain as those predictions that, if true, predict the truthfulness of all verifiable statements in the domain.
+We then show that experimental domains provide a natural Kolmogorov and second countable topology over their possibilities, while theoretical domains provide a natural $\sigma$-algebra.
 
@@
 During the course of the last century a few fields of science and engineering have developed an associated general theory.
-For example, in computer science theory one starts by defining what a computational device is and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input.
-In communications one defines what information is and shows that no encoding can outperform the Huffman coding.\footnote{Assuming a symbol-by-symbol coding with a known input probability distribution.} In control theory one defines what a system is and then looks for general strategies for control such as Kalman filtering.
-These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
+For example, in computer science theory one starts by defining what a computational device is\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input.\cite{Sipser} In communications one defines what information\cite{Shannon} is and shows that no encoding can outperform the Huffman coding.\cite{Pierce} In control theory one defines what a system is\cite{Brogan} and then looks for general strategies for control such as Kalman filtering.\cite{Kalman} These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
 That is: the results apply because of the mere definition of the subject matter.
@@
 No such general theory currently exists for physics, which is essentially composed of different theories loosely connected to each other, each with their own realm of applicability.
-However, we believe it can be created and in this work we present our basic strategy towards it.
-As one familiar to the general theories of other subjects would expect, we only need to focus on clarifying exactly what our starting points are and let the logic follow.
+However, we believe such a general theory can be created and in this work we present our basic strategy towards it.
+As one familiar with the general theories of other subjects would expect, we only need to focus on clarifying exactly what our starting points are and let the logic follow.
 That is: once we formalize exactly what it means to study a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories.
@@
 In this paper we briefly present the starting point for our general mathematical theory of experimental science.
-It is the product of a decade spent in reverse engineering the laws of particle mechanics (cite paper) and our subsequent work on axiomatizing the starting points (cite other paper?).
-Our overall program is large and ambitious and a journal article cannot treat it comprehensively.
+It is the product of a decade spent in reverse engineering the laws of particle mechanics\cite{Carc1} and our subsequent work on formalizing the starting points.\cite{Carc2} Our overall program is large and ambitious and a journal article cannot treat it comprehensively.
 Therefore we will only present an overview of the basic concepts, which are already well consolidated and provide a good template to understand our approach.
-We will only provide a sketch for the proofs, which can be found in our open access book (cite book) together with more details, discussions and insights.
+We will only provide a sketch for the proofs, which can be found in our open access book\cite{Carc3} together with more details, discussions and insights.
 
-The basic idea is that science deals with verifiable statements: assertion whose truth can be tested experimentally.
+The basic idea is that science deals with verifiable statements: assertions whose truth can be tested experimentally.
 We study the logic of these statements, which is different from the standard Boolean logic.
@@
 We will find that the topology is Kolmogorov and second countable, which limits the cardinality of the set of possibilities of a domain to the one of continuum.
-Note that these structure are indeed pervasive in science, and form the foundation for differential geometry (Riemannian and symplectic), measure theory, probability theory and many other mathematical tools already used in physics and other sciences.
+Note that these structures are indeed pervasive in science, and form the foundation for differential geometry (Riemannian and symplectic), measure theory, probability theory and many other mathematical tools already used in physics and other sciences.
 
@@
 This cannot work for us: what we are interested in is the meaning of the assertion and not how it is expressed.
-For example, \statement{this animal is a dog} and \statement{questo animale \`e un cane} are the same statement express in different languages.
+For example, \statement{this animal is a dog} and \statement{questo animale \`e un cane} represent the same fact expressed in different languages.
 As science is universal, it should not matter the language, units or reference system used to make an assertion.
-Therefore we introduce a variation of algebraic logic suited to our needs.
+While we will still use standard mathematical logic to run our formal system, we introduce a variation of algebraic logic to represent our ``informal'' assertions.
 We start by defining our truth bearer as:
@@
 Note how the first part of the definition captures the informal meaning of what we are describing, while the second part captures the part that is formalized.
-This pattern will be present in most of our definition and it serves to clarify both what is being formalized and how.
-Therefore, in science, the statement is an assertion while for the math is just an element in some set.
+This pattern will be present in most of our definitions and it serves to clarify both what is being formalized and how.
+Therefore, in science, the statement is an assertion while for the math it is just an element in some set.
 
 While in math the truth value is the focus of the logic system, in science it is generally established experimentally.
-But our model may constrain certain statements or statement combination to be ruled out.
+But our scientific model may constrain certain statements or statement combinations to be ruled out.
 For example, \statement{this animal is a dog} and \statement{this animal is a cat} can both be either true or false, but the statement \statement{this animal is a cat and a dog} can never be true.
-That is: the role of our logic system is not to keep track of what is true and false, but of what can't possibly be true and can't possibly be false.
-So we need to keep track of these relationships so that we can never have inconsistencies or paradoxes.
+That is: the role of our logic system is not to keep track of what is true and false, but of what cannot possibly be true and cannot possibly be false.
+We need to keep track of these relationships so that we can never have inconsistencies or paradoxes.
 
@@
 
-Note that the truth assignments are happening in our head: they are not actual physical entities.
+Note that the truth assignments are just hypotheticals.
+They are not actual physical entities.
 There is only one truth value, the one found experimentally.
-Logical consistency is defined on the truth assignments and not on the truth value.
+Logical consistency is defined on the truth assignments and not on the truth values.
 Therefore we need a way to track whether the meaning of each statement allows it to be true or not.
@@
 \begin{defn}
-	A \textbf{tautology} $\tautology$ is a statement that must be true simply because of its content. That is, $\possFn(\stmt) = \{\TRUE\}$.
+	A \textbf{tautology} $\tautology$ is a statement that must be true simply because of its content. That is, $\possFn(\tautology) = \{\TRUE\}$.
 \end{defn}
@@
 \begin{defn}
-	A \textbf{contradiction} $\contradiction$ is a statement that must be false simply because of its content. That is, $\possFn(\stmt) = \{\FALSE\}$.
+	A \textbf{contradiction} $\contradiction$ is a statement that must be false simply because of its content. That is, $\possFn(\contradiction) = \{\FALSE\}$.
 \end{defn}
@@
 	We can always construct a statement whose truth value arbitrarily depends on an arbitrary set of statements.
-Formally: given an arbitrary truth function $f_{\mathbb{B}} : \mathbb{B}^n \to \mathbb{B}$ there exists a function $f : \mathcal{S}^n \to \mathcal{S}$ such that
+Formally, given an arbitrary truth function $f_{\mathbb{B}} : \mathbb{B}^n \to \mathbb{B}$ there exists a function $f : \mathcal{S}^n \to \mathcal{S}$ such that
 	$$\truth(f(\stmt_1, ..., \stmt_n)) = f_{\mathbb{B}}(\truth(\stmt_1), ..., \truth(\stmt_n))$$
@@
 
-We will use the standard symbols $\NOT$, $\AND$ $\OR$ to indicate the negation (logical NOT), conjunction (logical AND) and disjunction (logical OR).
-With these three axioms, we can rederive all of of the rules of classical logic.
+We will use the standard symbols $\NOT$, $\AND$, $\OR$ to indicate the negation (logical NOT), conjunction (logical AND) and disjunction (logical OR).
+With these three axioms, we can rederive all of the rules of classical logic.
 First we define our equivalence.
@@
 
-This system not only allows to use formal classical logic while keeping the statements informal, but it also equipped to capture causal relationships.\footnote{Note that the axioms prevents any form of paradox, simply because the possibilities of a statement are never empty and a consistent truth assignment must exist.
-Therefore situations like \statement{this statement is not true} are ruled out simply because they cannot satisfy the axioms.} For example, $\stmt_1=$\statement{the thermometer indicator is between 24 C and 25 C} and $\stmt_s=$\statement{the temperature is between 24 C and 25 C} are equivalent because both $\stmt_1 \AND \NOT \stmt_2$ and $\NOT \stmt_1 \AND \stmt_2$ are contradictions (on the assumption that our thermometer is actually working).
+This system not only allows us to use formal classical logic while keeping the statements informal, but is also equipped to capture causal relationships.\footnote{Note that the axioms prevent any form of paradox, simply because the possibilities of a statement are never empty and a consistent truth assignment must exist.
+Therefore situations like \statement{this statement is not true} are ruled out simply because they cannot satisfy the axioms.} For example, $\stmt_1=$\statement{the thermometer indicator is between 24 C and 25 C} and $\stmt_2=$\statement{the temperature is between 24 C and 25 C} are equivalent because both $\stmt_1 \AND \NOT \stmt_2$ and $\NOT \stmt_1 \AND \stmt_2$ are contradictions (on the assumption that our thermometer is actually working).
 We can also define other semantic relationships.
@@
 \begin{defn}
-	Given two statement $\stmt_1$ and $\stmt_2$, we say that:
+	Given two statements $\stmt_1$ and $\stmt_2$, we say that:
 	\begin{itemize}
@@
 That is, $\possFn(f(S)) = f(\bigtimes\limits_{\stmt \in S} \possFn(\stmt))$ for any truth function $f : \mathbb{B}^{|S|} \to \mathbb{B}$.
-The negation of independence, will be noted by $\nindep$.
+The negation of independence will be noted by $\nindep$.
 \end{defn}
@@
 it is more specific), $\stmt_1$ $\ncomp$ $\stmt_3$ (i.e.
-they can't be true at the same time) and $\stmt_1$ $\indep$ $\stmt_4$ (i.e.
+they cannot be true at the same time) and $\stmt_1$ $\indep$ $\stmt_4$ (i.e.
 the truth of one tells us nothing about the truth of the other).
@@
 If we define statistical independence, it will have to be defined in a way that is compatible with statement independence.
-In the same way that set theory defines concepts common on all mathematics, the general theory defines basic concepts that are common to all scientific theories.
+In the same way that set theory defines concepts common across all mathematics, the general theory defines basic concepts that are common to all scientific theories.
 
@@
 
-We have the tools for dealing with assertions that are universal and non-contradictory, now we have to develop the tools for those that areevidence based as well.
+We have the tools for dealing with assertions that are universal and non-contradictory, now we have to develop the tools for those that are evidence based as well.
 
@@
 	\item find a swan
-	\item if it's black terminate successfully
+	\item if it is black terminate successfully
 	\item go to step 1
@@
 Because of non-termination, failure to verify is not verification of the negation.
-That is: the negation of a verifiable statement is not necessarily a verifiable statement.
+That is, the negation of a verifiable statement is not necessarily a verifiable statement.
 
@@
 But we cannot extend this to an infinite number of statements as we would never terminate.
-We can also verifying the disjunction of two statement: once one test terminates we are done.
+We can also verify the disjunction of two statements: once one test terminates we are done.
 And because we only need one test to terminate, we can generalize to a countable number of verifiable statements by following this procedure:
@@
 \end{enumerate}
-This procedure will run all tests for an arbitrary amount of time therefore, if one statement is true, it will make the test terminate successfully, even if all other tests would never terminate.
+This procedure will run all tests for an arbitrary amount of time.
+Therefore, if one statement is true, it will make the test terminate successfully, even if all other tests would never terminate.
 In light of this, we can set the following axioms.
@@
 In principle, given a set of verifiable statements we can simply start testing them one after the other.
-However, if we are given a set of uncountable statements than, even if we have an indefinitely long time at our disposal, we will not be able to create a procedure that eventually tests all statements.
+However, if we are given a set of uncountable statements then, even if we have an indefinitely long time at our disposal, we will not be able to create a procedure that eventually tests all statements.
 However, we may not need to actually run all tests for all statements.
@@
 
-Therefore it is the size of the basis, then, that matters and not the size of the set.
-If the basis is countable we can keep going and, if any statement is true, it will eventually be verified experimentally.\footnote{We assume we are given an indefinitely long time that because it's the upper limit, because it handles limits and because we haven't given constraints for how long or short a successful test can be (i.e.
-if ran a test for three days it may still successfully complete in three days and one minute).}
+Therefore it is the size of the basis that matters and not the size of the set.
+If the basis is countable we can keep going and, if any statement is true, it will eventually be verified experimentally.\footnote{We assume we are given an indefinitely long time because this handles limits and because we have not given constraints for how long or short a successful test can be (i.e.
+if we ran a test for three days without terminating, it may still successfully complete in three days and one minute).}
 
@@
 
-While verifiable statements define what can be scientifically studied, not all predictions are themselves verifiable.
+While verifiable statements define what can be scientifically studied, not all interesting scientific statements are directly verifiable.
 Consider the two statements \statement{there exists extra-terrestrial life} and \statement{there is no extra-terrestrial life}.
@@
 Yet, the second is still meaningful as a prediction: it predicts that the test for the first will never terminate.
-That is: while negations are not verifiable we can still logically talk about them when, for example, constructing truth assignments.
+That is, while negations are not verifiable we can still logically talk about them when, for example, constructing truth assignments.
 
@@
 As they provide all possible predictions, we focus on the ones that completely specify the truth value for all theoretical statements.
-For example, once we know that \statement{this animal is a cat} we know that \statement{this animal has whiskers}, that \statement{this animal has no feathers}and so on.
+For example, once we know that \statement{this animal is a cat} we know that \statement{this animal has whiskers}, that \statement{this animal has no feathers} and so on.
 
@@
 we learn how to test the DNA of animals) then the possibilities will change as well (e.g.
-the possible animal species is refined).
+the possible animal species are refined).
 We conclude this section with the following general result.
@@
 The proof is simply noting that each possibility can be labeled by the truth value of the countable basis of the experimental domain.
-We cannot have more possibilities than sequences of true/false and the set of all binary sequences has the cardinality of continuum.
+We cannot have more possibilities than sequences of true/false and the set of all binary sequences has the cardinality of the continuum.
 
@@
 
-Now that we have introduced the basic mathematical structures for our general theory, we show that their deep connection to other well established mathematical structures.
-
+Now that we have introduced the basic mathematical structures for our general theory, we show their deep connection to other well established mathematical structures.
 We first note that each verifiable statement can be written as the disjunction of a set of possibilities.
@@
 We define the map $U : \edomain \rightarrow 2^X$ that for each statement $\obs \in \edomain$ returns the set of possibilities compatible with it.
-That is: $U(\obs)\equiv\{ x \in X \, | \, x \comp \obs\}$.
+That is, $U(\obs)\equiv\{ x \in X \, | \, x \comp \obs\}$.
 We call $U(\obs)$ the \textbf{verifiable set} of possibilities associated with $\obs$.
@@
 
-The proof is a trivial application of the disjunction normal form of Boolean algebra.
+The proof is a trivial application of the disjunctive normal form of Boolean algebra.
 In fact, each possibility can be written as a minterm of a basis (i.e.
@@
 
-Since every statement is a set of possibilities, we can re-express the statement relationships in terms of sets according to this table:
+Since every statement is a set of possibilities, we can re-express the statement relationships in terms of set relationships according to this table:
 
@@
 Since the basis is countable, countable union is equivalent to arbitrary union.
-In other words: the set of all verifiable sets is a topology.
+In other words, the set of all verifiable sets is a topology.
 
@@
 It is second countable because a basis of the experimental domain corresponds to a sub-basis of the topology.
-
 One can also show that the natural topology is Hausdorff if and only if all possibilities are approximately verifiable (i.e.
 each possibility is the limit of a sequence of verifiable statements).
-The hope is that we can find physically meaningful definitions for all topological concepts.
+The hope is that we can find physically meaningful definitions for all relevant topological concepts.
 
@@
 
-This can be proven by showing that each possibility is equivalent to a class of Cauchy sequences, which corresponds to a real number.
+This can be proven by showing that each possibility is associated to an equivalence class of Cauchy sequences, which corresponds to a real number.
 Each open interval corresponds to a finite precision measurement.
 The possibilities are not themselves verifiable statements: we cannot measure a continuous quantity with infinite precision.
-But it is a theoretical statement: we can use the value to predict which finite precision measurement will be verified (the ones that contain that value).
+But it is a theoretical statement: we can use the value to predict which finite precision measurements will be verified (the ones that contain that value).
 
 This connection creates a strong bridge between the standard mathematical structures used in physics and their meaning.
-Every theorem, every proof now has a consistent and direct physical meaning.
-And given the foundational nature of topologies and $\sigma$-algebra, we have hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.
+Every theorem, every proof on those structures can now be given a direct physical meaning as well.
+And given the foundational nature of topologies and $\sigma$-algebras, we hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.
 
@@
 
-The general theory presented in this paper can be quickly summarized as follows.
+The general theory presented in this paper can be briefly summarized as follows.
 Science deals with statements that can be shown to be true experimentally.
@@
 the theoretical domain), and concentrate on those that fully predict all our measurements, the possibilities.
-By the very nature and properties of these objects, a topology and a $\sigma$-algebra is induced on the possibilities that retains the connection to what statements are verifiable and what statements constitute a valid prediction.
+By the very nature and properties of these objects, a topology and a $\sigma$-algebra is induced on the possibilities that retains the connection to which statements are verifiable and which statements constitute a valid prediction.
 
-The framework can be further be expanded by studying how additional assumptions constrain the verifiable statements.
-We have seen, for example, how assuming a quantity can be measured with arbitrary precision leads to the real numbers with standard topology.
+The framework can be further expanded by studying how additional assumptions constrain the verifiable statements.
+We have seen, for example, how assuming a quantity can be measured with arbitrary precision leads to the real numbers with the standard topology.
 Similarly, other assumptions will lead to other mathematical structures.
-This expansion is the aim of current and future work.
+This expansion is the aim of future work.
 
@@
 
-The other benefit is that requires no interpretation.
-All mathematical definition are capturing a well defined physical concept.
+The other benefit is that no interpretation is required.
+All mathematical definitions are capturing well defined physical concepts.
 Therefore every theorem, every step in every proof will also have a well defined physical meaning.
-Since we know what we have put in, we know what comes out.
 
-We feel that this type of work is of significance importance for the scientific community in general and for the physics community in particular if we are ever to truly understand what scientific theories are and what are they meant to be describing.
+We feel that this type of work is of significance for the scientific community in general and for the physics community in particular if we are ever to truly understand what scientific theories are meant to be describing.
 
@@
 
+\begin{thebibliography}{1}
+
+	\bibitem{Turing}
+	A. M. Turing: On Computable Numbers, with an Application to the Entscheidungsproblem". Proceedings of the London Mathematical Society 2, 42: 230–265, doi:10.1112/plms/s2-42.1.230, 1936.
+
+	\bibitem{Sipser} M. Sipser: An Introduction to the Theory of Computation, 3rd edition. Cengage Learning, Boston, USA, 2013.
+
+	\bibitem{Shannon} C.E.Shannon: A Mathematical Theory of Communication. The Bell System Technical Journal,
+	Vol. 27, pp. 379–423, 623–656, July, October, 1948.
+
+	\bibitem{Pierce} J.R. Pierce: An Introduction To Information Theory: Symbols, Signals and Noise. Dover Books on Mathematics, Second Edition, 1980.
+
+	\bibitem{Brogan} W. L. Brogan: Modern Control Theory, 3rd edition. Pearson, 1990.
+
+	\bibitem{Kalman} R. E. Kalman: A new approach to linear filtering and prediction problems. Journal of Basic Engineering. 82 (1): 35–45, doi:10.1115/1.3662552, 1960.
+
+	\bibitem{Carc1} G.
+Carcassi, C.
+A.
+Aidala, D.
+J.
+Baker and L.
+Bieri: From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics.
+Journal of Physics Communications, 2, 4, 045026, 2018.
+
+	\bibitem{Carc2} C. A. Aidala, G. Carcassi, M. J. Greenfield: Topology and experimental distinguishability. arXiv:1708.05492
+
+	\bibitem{Carc3} G. Carcassi, C. A. Aidala: Assumptions of Physics, v0.1. http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.
+\end{thebibliography}
+
 \end{document}

### 2018-05-14 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Fix citations

@@
 During the course of the last century a few fields of science and engineering have developed an associated general theory.
-For example, in computer science theory one starts by defining what a computational device is\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input.\cite{Sipser} In communications one defines what information\cite{Shannon} is and shows that no encoding can outperform the Huffman coding.\cite{Pierce} In control theory one defines what a system is\cite{Brogan} and then looks for general strategies for control such as Kalman filtering.\cite{Kalman} These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
+For example, in computer science theory one starts by defining what a computational device is\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input.\cite{Sipser} In communications one defines what information is\cite{Shannon} and shows that no encoding can outperform the Huffman coding.\cite{Pierce} In control theory one defines what a system is\cite{Brogan} and then looks for general strategies for control such as Kalman filtering.\cite{Kalman} These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
 That is: the results apply because of the mere definition of the subject matter.
@@
 
-	\bibitem{Shannon} C.E.Shannon: A Mathematical Theory of Communication. The Bell System Technical Journal,
+	\bibitem{Shannon} C. E. Shannon: A Mathematical Theory of Communication. The Bell System Technical Journal,
 	Vol. 27, pp. 379–423, 623–656, July, October, 1948.
 
-	\bibitem{Pierce} J.R. Pierce: An Introduction To Information Theory: Symbols, Signals and Noise. Dover Books on Mathematics, Second Edition, 1980.
+	\bibitem{Pierce} J. R. Pierce: An Introduction To Information Theory: Symbols, Signals and Noise. Dover Books on Mathematics, Second Edition, 1980.
 

### 2018-06-14 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Changed discussion

@@
 Our program is altogether different and orthogonal.
-Our subject matter are the scientific models themselves.
+Our subject matter is the scientific models themselves.
 Our aim is to understand what aspects of a scientific theory are already constrained by the mere fact that it is a scientific theory or by the assumptions one bakes into them.
@@
 
-To give an example, consider an experimental domain formed by statements like \statement{this quantity is between $a$ and $b$}.
+\section{Brief discussion}
+
+The connection we outlined creates a strong bridge between the standard mathematical structures used in physics and their meaning.
+Every theorem, every proof on those structures can now be given a direct physical meaning as well.
+And given the foundational nature of topologies and $\sigma$-algebras, we hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory\footnote{Probability statements will be of the form \statement{The position of the center of mass of the system is between 12 and 13 nm, 20\% to 21\% of the time.}} and so on.
+
+To give an example, consider an experimental domain where the basis is composed of statements like \statement{this quantity is between $a$ and $b$} where $a$ and $b$ are different values that can be arbitrarily close.
 For example, \statement{the distance between the earth and the moon is between 384 and 385 thousand Km}.
-And suppose that the precision can be arbitrarily narrowed.
-We can define the following.
+Each verifiable statement of the basis corresponds to an open interval of the real line, therefore we find a correspondence between arbitrary precision measurements and the standard topology on the reals, as this is the one generated by open intervals.
+Note, in fact, that this topology is second-countable and Kolmogorov.
 
-\begin{defn}
-	An experimental domain $\edomain_X$ is said to allow \textbf{arbitrary precision} if it describes a quantity measurable with arbitrary precision.
-Formally, there exists a countable basis $\basis \subseteq \edomain_X$ such that:
-	\begin{itemize}
-		\item we can define a function $\operatorname{bounds} : \basis \to \mathbb{Q} \times \mathbb{Q}$ such that $\operatorname{bounds}(\stmt) = (a, b)$ is a tuple where $a < b$ for all $\stmt \in \basis$
-		\item for every pair of rational numbers $(a, b) \in \mathbb{Q} \times \mathbb{Q}$ there exists an $\obs \in \basis$ such that $\operatorname{bounds}(\stmt) = (a, b)$
-		\item if $\stmt = \stmt_1 \AND \stmt_2$, $\operatorname{bounds}(\stmt_1) = (a_1, b_1)$ and $\operatorname{bounds}(\stmt_2) = (a_2, b_2)$, then $\stmt \equiv \contradiction$ if $b_1 \leq a_2$ or $b_2 \leq a_1$ otherwise $\stmt \nequiv \contradiction$ and $\operatorname{bounds}(\stmt) = (\max(a_1, a_2), \min(b_1, b_2))$
-	\end{itemize}
-\end{defn}
+The possibilities (i.e.
+\statement{the mass of the photon is precisely zero}) are not themselves verifiable since infinite precision measurements of a continuous quantity are not possible.
+Yet, their negation (i.e.
+\statement{the mass of the photon isn't precisely zero}) could be verified in practice.
+This is expressed mathematically by the fact that singletons are closed sets.
+The theoretical domain, instead, corresponds to the standard Borel algebra.
 
-Then it can be shown that:
-
-\begin{thrm}[Arbitrary precision is continuity]\label{thrm_arbitrary_precision_is_continuity}
-	Given an experimental domain $\edomain_X$, the set of possibilities with their natural topology is homeomorphic to the reals with the standard topology if and only if the domain allows arbitrary precision.
-\end{thrm}
-
-This can be proven by showing that each possibility is associated to an equivalence class of Cauchy sequences, which corresponds to a real number.
-Each open interval corresponds to a finite precision measurement.
-The possibilities are not themselves verifiable statements: we cannot measure a continuous quantity with infinite precision.
-But it is a theoretical statement: we can use the value to predict which finite precision measurements will be verified (the ones that contain that value).
-
-This connection creates a strong bridge between the standard mathematical structures used in physics and their meaning.
-Every theorem, every proof on those structures can now be given a direct physical meaning as well.
-And given the foundational nature of topologies and $\sigma$-algebras, we hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.
+Similarly, one can imagine verifiable statements for relationships (i.e.
+\statement{when the temperature of the mercury column is between 24 and 25 C, its height is between 24 and 25 mm}) and for statistical variables (i.e.
+\statement{if this coin is tossed enough times, the ratio of heads will be between 45\% and 55\%}).
+Ultimately the general theory will need to show precisely what mathematical structures map to experimental domains formed by statements of these types and under what assumptions.
 

### 2018-06-14 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Adding acknowledgements

@@
 Therefore we will only present an overview of the basic concepts, which are already well consolidated and provide a good template to understand our approach.
-We will only provide a sketch for the proofs, which can be found in our open access book\cite{Carc3} together with more details, discussions and insights.
+We will only provide a sketch for the proofs, which can be found in the draft of our open access book\cite{Carc3} together with more details, discussions and insights.
 
@@
 
+\section{Acknowledgments}
+
+We thank Mark J.
+Greenfield for the thorough review of the mathematical framework.
+We thank Josh Hunt for comments on the philosophical underpinning of this work.
+We would also like to acknowledge the numerous individuals who commented on the drafts of this paper and our previous works leading up to it.
+
 \bibliography{bibliography}
@@
 
-	\bibitem{Carc2} C. A. Aidala, G. Carcassi, M. J. Greenfield: Topology and experimental distinguishability. arXiv:1708.05492
+	\bibitem{Carc2} C. A. Aidala, G. Carcassi, M. J. Greenfield: Topology and experimental distinguishability. arXiv:1708.05492, 2017.
 
-	\bibitem{Carc3} G. Carcassi, C. A. Aidala: Assumptions of Physics, v0.1. http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.
+	\bibitem{Carc3} G. Carcassi, C. A. Aidala: Assumptions of Physics (in preparation). http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.
 \end{thebibliography}

### 2018-06-22 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Final review before submission

@@
 \begin{abstract}
-	In this article we lay the ground work for a general mathematical theory of experimental science.
+	In this article we lay the groundwork for a general mathematical theory of experimental science.
 The starting point will be the notion of verifiable statements, those assertions that can be shown to be true with an experimental test.
-We study the algebra of such objects and show how it is closed only on finite conjunction and countable disjunction.
+We study the algebra of such objects and show how it is closed only under finite conjunction and countable disjunction.
 With simple constructions, we show that the set of possible cases distinguishable by verifiable statements is equipped with a natural Kolmogorov and second countable topology and a natural $\sigma$-algebra.
@@
 During the course of the last century a few fields of science and engineering have developed an associated general theory.
-For example, in computer science theory one starts by defining what a computational device is\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input.\cite{Sipser} In communications one defines what information is\cite{Shannon} and shows that no encoding can outperform the Huffman coding.\cite{Pierce} In control theory one defines what a system is\cite{Brogan} and then looks for general strategies for control such as Kalman filtering.\cite{Kalman} These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
-That is: the results apply because of the mere definition of the subject matter.
+For example, in computer science theory one starts by defining what a computational device is~\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input~\cite{Sipser}.
+In communications one defines what information is~\cite{Shannon} and shows that no encoding can outperform the Huffman coding~\cite{Pierce}.
+In control theory one defines what a system is~\cite{Brogan} and then looks for general strategies for control such as Kalman filtering~\cite{Kalman}.
+These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
+That is, the results apply because of the mere definition of the subject matter.
 
@@
 As one familiar with the general theories of other subjects would expect, we only need to focus on clarifying exactly what our starting points are and let the logic follow.
-That is: once we formalize exactly what it means to study a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories.
+That is, once we formalize exactly what it means to study a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories.
 For example, we will see that any set of objects that are physically distinguishable must have cardinality no greater than continuum.
@@
 Our subject matter is the scientific models themselves.
-Our aim is to understand what aspects of a scientific theory are already constrained by the mere fact that it is a scientific theory or by the assumptions one bakes into them.
+Our aim is to understand what aspects of a scientific theory are already constrained by the mere fact that it is a scientific theory or by the assumptions one builds into them.
 While this general theory may be of some utility to someone working on a grand-unified theory as it may help narrow the search, neither one will lead to the other.
@@
 In this paper we briefly present the starting point for our general mathematical theory of experimental science.
-It is the product of a decade spent in reverse engineering the laws of particle mechanics\cite{Carc1} and our subsequent work on formalizing the starting points.\cite{Carc2} Our overall program is large and ambitious and a journal article cannot treat it comprehensively.
+It is the product of a decade spent in reverse engineering the laws of particle mechanics~\cite{Carc1} and our subsequent work on formalizing the starting points~\cite{Carc2}.
+Our overall program is large and ambitious and a journal article cannot treat it comprehensively.
 Therefore we will only present an overview of the basic concepts, which are already well consolidated and provide a good template to understand our approach.
-We will only provide a sketch for the proofs, which can be found in the draft of our open access book\cite{Carc3} together with more details, discussions and insights.
+We will only provide a sketch for the proofs, which can be found in the draft of our open access book~\cite{Carc3} together with more details, discussions and insights.
 
@@
 
-The first thing we need to do is to characterize what science is and what its scope is. We do so by introduction the following:
+The first thing we need to do is to characterize what science is and what its scope is. We do so by introducing the following:
 
@@
 As science is universal, it should not matter the language, units or reference system used to make an assertion.
-While we will still use standard mathematical logic to run our formal system, we introduce a variation of algebraic logic to represent our ``informal'' assertions.
+While we will still use standard mathematical logic for the formal system, we introduce a variation of algebraic logic to represent our ``informal'' assertions.
 We start by defining our truth bearer as:
@@
 But our scientific model may constrain certain statements or statement combinations to be ruled out.
-For example, \statement{this animal is a dog} and \statement{this animal is a cat} can both be either true or false, but the statement \statement{this animal is a cat and a dog} can never be true.
-That is: the role of our logic system is not to keep track of what is true and false, but of what cannot possibly be true and cannot possibly be false.
+For example, the statements \statement{this animal is a dog} and \statement{this animal is a cat} can both be either true or false, but the statement \statement{this animal is a cat and a dog} can never be true.
+That is, the role of our logic system is not to keep track of what is true and false, but of what cannot possibly be true and cannot possibly be false.
 We need to keep track of these relationships so that we can never have inconsistencies or paradoxes.
@@
 This means that we are never going to be able to experimentally distinguish between elements of greater cardinality.
-The set of discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$, for example, has greater cardinality and therefore it will never be associated with any physically distinguishable concept.
-All the issues with large cardinals are not something science will ever be interested in.
+The set of topologically discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$, for example, has greater cardinality and therefore it will never be associated with any physically distinguishable concept.\footnote{However, discontinuous function with up to countably many discontinuities can be seen as the limit of a sequence of continuous functions and therefore may be used as a convenient approximation of a physical concept.} All the issues with large cardinals are not something science will ever be interested in.
 It does not matter what system we are describing, what experimental techniques we are using or how clever we are.
@@
 Every theorem, every proof on those structures can now be given a direct physical meaning as well.
-And given the foundational nature of topologies and $\sigma$-algebras, we hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory\footnote{Probability statements will be of the form \statement{The position of the center of mass of the system is between 12 and 13 nm, 20\% to 21\% of the time.}} and so on.
+And given the foundational nature of topologies and $\sigma$-algebras, we hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.
 
@@
 
-The framework can be further expanded by studying how additional assumptions constrain the verifiable statements.
-We have seen, for example, how assuming a quantity can be measured with arbitrary precision leads to the real numbers with the standard topology.
+The framework will be further expanded by studying how additional assumptions constrain the verifiable statements.
+Assuming a quantity can be measured with arbitrary precision will lead to the real numbers with the standard topology.
 Similarly, other assumptions will lead to other mathematical structures.
@@
 One benefit of the general theory is to bring to the forefront our hidden assumptions.
-We have seen that real numbers are associated with arbitrary precision.
-What happens in the case of limited precision, when we cannot assume precision can always be increased indefinitely?
+If real numbers are associated with arbitrary precision, what happens in the case of limited precision, when we cannot assume precision can always be increased indefinitely?
 Would we still have real numbers?

### 2018-09-15 Articles/LogicalTopologicalFramework/article.tex
Added first draft
Confused by github but learning new things is good!

\documentclass[psamsfonts]{amsart}

%-------Packages---------
\usepackage{amssymb,amsfonts}
\usepackage{url}
\usepackage[all,arc]{xy}
\usepackage{enumerate}
\usepackage{mathrsfs}
\usepackage{tikz-cd}

%--------Theorem Environments--------
%theoremstyle{plain} --- default
\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{quest}[thm]{Question}
\newtheorem{fact}[thm]{Fact}
\newtheorem{axiom}[thm]{Axiom}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem{defns}[thm]{Definitions}
\newtheorem{con}[thm]{Construction}
\newtheorem{exmp}[thm]{Example}
\newtheorem{exmps}[thm]{Examples}
\newtheorem{notn}[thm]{Notation}
\newtheorem{notns}[thm]{Notations}
\newtheorem{addm}[thm]{Addendum}
\newtheorem{exer}[thm]{Exercise}

\theoremstyle{remark}
\newtheorem{rem}[thm]{Remark}
\newtheorem{rems}[thm]{Remarks}
\newtheorem{warn}[thm]{Warning}
\newtheorem{sch}[thm]{Scholium}

\makeatletter
\let\c@equation\c@thm
\makeatother
\numberwithin{equation}{section}

\bibliographystyle{plain}

%--------New Commands-------
\newcommand{\R}{\mathbb{R}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\C}{\mathbb{C}}

\renewcommand\Re{\operatorname{Re}}
\renewcommand\Im{\operatorname{Im}}
\newcommand{\aand}{\text{ and }}

% Adds double bracket symbols
\usepackage{stmaryrd}

% Latex symbol guide at http://mirrors.ibiblio.org/CTAN/info/symbols/comprehensive/symbols-letter.pdf

% LOGIC symbols
% -------------

% Allows to create negation symbols
\usepackage{MnSymbol}

\DeclareMathOperator{\truth}{truth}
\DeclareMathOperator{\possFn}{poss}
\DeclareMathOperator{\result}{result}
\DeclareMathOperator{\idFn}{id}

\def\TRUE{\textsc{true}}
\def\FALSE{\textsc{false}}

\def\SUCCESS{\textsc{success}}
\def\FAILURE{\textsc{failure}}
\def\UNDEF{\textsc{undefined}}

% Symbols for statements set
\def\stmtSet{\mathcal{S}}
\def\vstmtSet{\mathcal{S}_\textsf{v}}
\def\dstmtSet{\mathcal{S}_\textsf{d}}

% Symbols for tautology and contradiction
\def\tautology{\top}
\def\contradiction{\bot}

% Symbols for "compatibility" and "incompatibility"
\def\comp{\doublefrown}
\def\ncomp{\ndoublefrown}

% Symbols for "narrower" and "wider"
\def\narrower{\preccurlyeq}
\def\nnarrower{\npreccurlyeq}
\def\snarrower{\prec}
\def\nsnarrower{\nprec}
\def\broader{\succcurlyeq}
\def\nbroader{\nsucccurlyeq}
\def\sbroader{\succ}
\def\nsbroader{\nsucc}

% Symbol for "independent" and "correlated"
\def\indep{\upmodels}
\def\nindep{\nupmodels}

% Aliases for logical operations
\def\AND{\wedge}
\def\bigAND{\bigwedge}
\def\OR{\vee}
\def\bigOR{\bigvee}
\def\NOT{\neg}

% Formatting for statements
\newcommand{\stmt}[1][s] {\mathsf{#1}}
% Formatting for experimental tests
\newcommand{\expt}[1][e] {\mathsf{#1}}
\newcommand{\exptSet}{\mathcal{E}}
% Formatting for observations
\newcommand{\obs}[1][s] {\mathsf{#1}}
% Formatting for possibilities
\newcommand{\resPoss}[1][x] {\mathring{#1}}
\newcommand{\estPoss}[1][x] {\dot{#1}}

% Formatting for experimental domain
\newcommand{\edomain}[1][D] {\mathcal{#1}}

% Formatting for theoretical domain
\newcommand{\tdomain}[1][D] {\bar{\mathcal{#1}}}

\newcommand{\basis}[1][B] {\mathcal{#1}} % Basis

% Formatting for experimental relationships
\newcommand{\erel}[1][r] {#1}

% Formatting for sentence statements
\newcommand{\statement}[1] {\emph{``#1"}}

% Formatting for reference
\newcommand{\refStmt}[1][r]{\textbf{#1}}

%--------Meta Data: Fill in your info------
\title{A logical and topological framework for experimental science}

\author{Christine Aidala, Gabriele Carcassi, and Mark Greenfield}

\date{}

\begin{document}
\begin{abstract}
In this paper we construct a logical and topological framework which encompasses a mathematical theory of the experimental science. Beginning with an axiomatic notion of a (scientifically) verifiable statement, we build an algebra of statements with a natural topology. In particular, we show that any collection of objects distinguishable by scientific methods has a second-countable topology with $\mathsf{T}_0$ separability, and those which can be approximated with high precision have Hausdorff topologies. The major definitions and theorems have straightforward physical meaning as well as precise mathematical underpinnings. This paper is intended to give a concise introduction and foundation to a growing body of work. 
\end{abstract}

\maketitle

\section{Introduction}
Abstract mathematical frameworks have proven to be extremely useful in various areas of science and engineering, enabling one to derive far-reaching consequences in a systematic manner. However, there does not exist a mathematical theory underlying experimental science in general. Many questions both general and specific may be made more tractable by such a theory. For example, understanding which kinds of statements it even makes sense to test experimentally, or building a general method for classifying which possible outcomes are nearer to each other are both problems which are usually dealt with on a case-by-case basis. For the field of particle mechanics, this effort is well underway with \cite{carc1}. Virtually all scientific tests for quantitative data have finite precision, which motivates the deep and pervasive relationship science has with topology. We might think of open sets as those for which we can test if a certain value lies within them, for example.

In this work, we will develop a logical framework which can be applied to experimental sciences in general. We take as a general assumption that science is universal and non-contradictory. We begin by defining and axiomatizing \emph{verifiable statements}, which essentially are those statements which it is possible (given arbitrarily many resources) to test experimentally. The collection of these statements will be endowed with the structure of a Boolean algebra. Following this, we will define \emph{possibilities} of an \emph{experimental domain}, which capture the notion of all possible outcomes to any sequence of tests. Using only well-justified axioms and standard proof methods, we will show the following: 
\begin{thm}\label{cardinality_thm}
If $X$ is a set of possibilities for an experimental domain, then $X$ has cardinality at most that of the continuum.
\end{thm}
\begin{thm}
\label{main}
If $X$ is a set of possibilities for an experimental domain, then $X$ has a natural topology built from verifiable sets which is second-countable and has $T_0$ separability. The possibilities can verified with arbitrarily high precision if and only if it is also Hausdorff. 
\end{thm}
The topology is called the \emph{topology of experimental possibility}. These will all be restated with more precise terminology in later sections. Along the way, we will provide physical meaning to many other topological and logical concepts in this context.

These ideas form the foundation of a growing body of work. The continually evolving manuscript \cite{carc3} contains full coverage of the ideas of this paper and many additional topics. The preprint \cite{carc2} contains some precursor ideas to the content of the present paper. The ideas in Theorem \ref{main} can be used to understand experimental relationships between different experimental domains using continuous maps and homeomorphisms. This will be covered in a later paper. Many other ideas in development are based on the contents of the present paper.

This paper is organized as follows. In Section \ref{statements} we define statements and their Boolean structure, as well as a partial ordering by precision. Following this is Section \ref{verifiable} in which we pick out verifiable statements and special collections of statements which we call experimental domains. In Section \ref{possibilities} we will define the possibilities of a domain and prove Theorem \ref{cardinality_thm}. Finally, in Section \ref{topology} using the space of possibilities we build a topological space and prove Theorem \ref{main} as a conglomeration of several smaller propositions. We close in Section \ref{conclusion} with a few remarks and questions.

\noindent {\bf Acknowledgements:} We thank Josh Hunt for comments on the philosophical ideas of this work, as well as various others who have provided comments and suggestions along the way. The third author is supported by the National Science Foundation Graduate Research Fellowship Program under Grant No. DGE\#1256260. This project is partially supported by the MCubed program at the University of Michigan.

\section{Statements}
\label{statements}

Our most basic objects will be statements. These will not be defined as sentences in a formal language; rather, we will give several axioms which characterize them as abstract objects. The physical meaning of each axiom will be straightforward, and one can see how scientific statements in practice are captured by our abstract statements.

The main reason we cannot use traditional logic and formal languages is that we need to distinguish the \emph{truth value} of a statement from its \emph{content}. In science, one does not a priori know (or even have access to) the truth value of all statements, but one can interpret the content to understand their relationships with other statements. For example, the statements ``the electron has mass less than $10^{-30}kg$" and ``the electron has mass greater than $10^{-30}kg$" cannot both be true simultaneously, and we can determine this without knowing the actual truth values. Moreover, we wish to understand logical implication with respect to content as separate from standard conditionals based on the truth value alone. Our axioms will capture the content and truth values as separate concepts. In what follows, we denote the set of truth values by $\mathbb{B} = \{\TRUE,\FALSE\}$.

\subsection{Building the collection of statements}

\begin{defn}
A \emph{statement} is an object with a truth value, ascertained by a truth function outputting a Boolean, and logical content, defined by satisfying the Consistency Axiom (\ref{consistent}) and the Completeness Axiom (\ref{completeness}). The set of all statements is denoted $\mathcal{S}$. The truth function is denoted $\truth:\mathcal{S}\to\mathbb{B}$.
\end{defn}

\begin{axiom}[Consistency Axiom]
\label{consistent}
Given a collection of statements $\{\stmt_{\lambda}\}_{\lambda\in\Lambda}$ for some (possibly uncountable) index set $\Lambda$, and given a set of truth values $\{t_{\lambda}\}_{\lambda\in\Lambda}$, where each $t_{\lambda}\in\mathbb{B}$, one can ascertain whether simultaneously supposing $\truth(\stmt_{\lambda}) = t_{\lambda}$ for all $\lambda\in\Lambda$ is a logically \emph{consistent truth assignment}. This is independent of the actual truth values of the statements. 
\end{axiom}

This axiom captures the general notion of content of an individual statement. Our earlier example of two statements about electron mass cannot both be true, but all other combinations (both false, one true and one false) are consistent truth assignments. We next define the possibilities of a statement in terms of consistent truth values of single statements.

\begin{defn}
The \emph{possibilities} of a statement $\stmt\in\mathcal{S}$ are the possible truth values allowed by the content of the statement. We define the possibilities function as
$$
\possFn:\mathcal{S}\to\{\{\FALSE,\TRUE\},\{\TRUE\},\{\FALSE\}\}
$$
where $\possFn(\stmt)$ is defined to be the set $S\subseteq\mathbb{B}$ of truth values such that $\{\stmt\},\{B\}$ is a consistent truth assignment for each $B\in S$. 
\end{defn}

\begin{defn}
A \emph{tautology} is a statement whose possibilities are $\{\TRUE\}$. A \emph{contradiction} is a statement whose possibilities are $\{\FALSE\}$.
\end{defn}

The notion of possibilities of a statement captures the difference between statements whose truth values can be obtained a priori (such as mathematical statements or tautologies), and those which must be experimentally tested. It is immediate that $\truth(\stmt)\in\possFn(\stmt)$ for all statements $\stmt$. In fact, for any collection of statements, assigning them their truth values is always a consistent truth assignment, since we assume true scientific statements to be non-contradictory. However, there are many consistent truth assignments which do not arise from the truth values of the included statements. Conversely, for any consistent truth assignment, each truth value $t_{\lambda}$ must be in $\possFn(\stmt_{\lambda})$. We now can define the Completeness Axiom:

\begin{axiom}[Completeness Axiom]
\label{completeness}
Given a collection of statements $\{\stmt_{\lambda}\}_{\lambda\in\Lambda}$ for some (possibly uncountable) index set $\Lambda$, let $A\subseteq\Lambda$, and let $\{t_a\}_{a\in A}$ be a consistent truth assignment for $\{\stmt_a\}_{a\in A}$. Then there exists a consistent truth assignment $\{t_{\lambda}\}_{\lambda\in\Lambda}$ for $\{\stmt_{\lambda}\}_{\lambda\in\Lambda}$ such that $t_{\lambda} = t_a$ for all $\lambda\in A$. 
\end{axiom}
This axiom expands the notion of what is considered to be a possible truth value. No possible truth value of any statement can lead to a paradox in larger collections of statements. These two axioms define our basic collection of statements. Next, we will give an additional axiom which will allow us to create new statements from old ones.

\begin{axiom}[Construction axiom]
Let $f_{\mathbb{B}}$ be a Boolean function $f_{\mathbb{B}}:\mathbb{B}^{\Lambda}\to\mathbb{B}$, where $\Lambda$ is any cardinal. Then there exists a function on the set of statements $f:(\mathcal{S}^*)^{\Lambda}\to\mathcal{S}$ such that
		$$\truth(f(\stmt_1, ..., \stmt_n)) = f_{\mathbb{B}}(\truth(\stmt_1), ..., \truth(\stmt_n))$$
with the same relationship holding for every consistent truth assignment. 
\end{axiom}

The point of the Construction axiom is to extend our usual Boolean logical operations to statements. We will now define them in this context.

\begin{notn}
The usual Boolean functions AND, OR, NOT, denoted respectively by $\AND,\ \OR,\ \NOT$, will also refer to the corresponding operations on statements. They will be applied to arbitrary collections of statements. The Construction axiom ensures that $\mathcal{S}$ will contain the prescribed statements. 
\end{notn}

It is important to note that the possibilities of a statement composed from other statements cannot be determined just by looking at the truth values and possibilities of individual statements alone. For example, two statements $\stmt_1$ and $\stmt_2$ which both have $\mathbb{B}$ as their possibilities may have any subset of $\mathbb{B}$ as the possibilities of their conjunction $\stmt_1\AND\stmt_2$ (regardless of the truth values).

Now, we will define an equivalence relation on the set of statements which will dramatically reduce the collection of statements to a more manageable size.

\begin{defn}
Two statements $\stmt_1$ and $\stmt_2$ are \emph{equivalent}, denoted $\stmt_1 \equiv \stmt_2$, if and only if $$\stmt_0:=(\stmt_1 \AND \stmt_2) \OR (\NOT\stmt_1 \AND \NOT\stmt_2)$$ has $\possFn(\stmt_0)=\{\TRUE\}$. 
\end{defn}

Informally, this means that $\stmt_1\equiv\stmt_2$ if and only if they must \emph{a priori} share the same truth value, independent of the actual truth values. It is immediate that two equivalent statements must have the same truth value and possibilities, since in any consistent truth assignment for $\{\stmt_1,\stmt_2,\stmt_0\}$ (given in the definition above), we must have that the statement $\stmt_0$ true. We will first confirm that this is an equivalence relation:

\begin{prop}
Statement equivalence is an equivalence relation.
\end{prop}
\begin{proof}
Reflexivity and symmetry are clear. For transitivity, let $\stmt_i$, $i=1,2,3$ be statements with $\stmt_1\equiv\stmt_2$ and $\stmt_2\equiv\stmt_3$. Consider the collection $\{\stmt_1,\stmt_2,\stmt_3\}$. Every consistent truth assignment must give the same truth value to $\stmt_1$ and $\stmt_2$, as well as $\stmt_2$ and $\stmt_3$, since these pairs are equivalent. It follows that $\stmt_1\equiv\stmt_3$. 
\end{proof}

An alternative way to view statement equivalence is as follows:

\begin{prop}\label{prop_equivalent_is_iff}
	Let $\stmt_1, \stmt_2 \in \mathcal{S}$ be two statements. Then $\stmt_1\equiv\stmt_2$ if and only if in all consistent truth assignments, $\stmt_1$ and $\stmt_2$ must have the same truth value.
\end{prop}

\begin{proof}
Suppose all consistent truth assignments require $\stmt_1$ and $\stmt_2$ to share a truth value. Consider a consistent truth assignment $\{t_1, t_2, t_3\}$ for the sequence of statements $$\{\stmt_1, \stmt_2, (\stmt_1 \AND \stmt_2) \OR (\NOT\stmt_1 \AND \NOT\stmt_2)\}.$$ By construction, a consistent truth assignment must then either be $\{\TRUE, \TRUE, \TRUE\}$ or $\{\FALSE, \FALSE, \TRUE\}$. In all cases $t_3=\TRUE$, so $$\possFn((\stmt_1 \AND \stmt_2) \OR (\NOT\stmt_1 \AND \NOT\stmt_2)) = \{\TRUE\}.$$ Therefore $\stmt_1 \equiv \stmt_2$.

For the converse, recall that $\possFn((\stmt_1 \AND \stmt_2) \OR (\NOT\stmt_1 \AND \NOT\stmt_2)) = \{\TRUE\}$. Denote this statement by $\stmt_0$. Because any consistent truth assignment for $\{\stmt_1,\stmt_2\}$ can be extended to a consistent truth assignment for $\{\stmt_1,\stmt_2,\stmt_0\}$, and $\possFn(\stmt_0)=\{\TRUE\}$, the claim follows. 
\end{proof}

It is an immediate consequence of the definition that all tautologies are equivalent, and all contradictions are equivalent. We will interchangeably denote tautologies and their equivalence class by $\tautology$, and similarly contradictions and their equivalence class will be denoted by $\contradiction$. Proposition \ref{prop_equivalent_is_iff} allows one to quickly see that if two Boolean functions are equal, then the corresponding functions on $\mathcal{S}$ must give equivalent results. We then arrive at the following:

\begin{cor}\label{boolean_properties}
		The set of all statements $\mathcal{S}$ satisfies the following properties, where $a,b,c\in\mathcal{S}$, and $\tautology$ and $\contradiction$ denote any tautology and contradiction, respectively:
		\begin{itemize}
			\item associativity: $a \OR (b \OR c) \equiv (a \OR b) \OR c$, $a \AND (b \AND c) \equiv (a \AND b) \AND c$
			\item commutativity: $a \OR b \equiv b \OR a$, $a \AND b \equiv b \AND a$
			\item absorption: $a \OR (a \AND b) \equiv a$, $a \AND (a \OR b) \equiv a$
			\item identity: $a \OR \contradiction \equiv a
			$, $a \AND \tautology \equiv a$
			\item distributivity: $a \OR (b \AND c) \equiv (a \OR b) \AND (a \OR c)$, $a \AND (b \OR c) \equiv (a \AND b) \OR (a \AND c)$
			\item complements: $a \OR \NOT a \equiv \tautology$, $a \AND \NOT a \equiv \contradiction$
			\item De Morgan: $\NOT a \OR \NOT b \equiv \NOT (a \AND b)$, $\NOT a \AND \NOT b \equiv \NOT (a \OR b)$
		\end{itemize}
		In particular, the set of equivalence classes of $\mathcal{S}$ is a Boolean algebra. These properties hold for infinite collections as well. 
	\end{cor}
\begin{proof}
The left-hand side of each equation is equivalent as a Boolean function to the right-hand side. Proposition \ref{prop_equivalent_is_iff} then implies the left- and right-hand sides must give equivalent statements. 
\end{proof}

While $\mathcal{S}$ is not technically a Boolean algebra itself (since e.g. there are many nonequal, but equivalent, tautologies and contradictions), it will enjoy many of the useful consequences of being a Boolean algebra when using statement equivalence rather than equality. We call the operations and properties defined thus far the \emph{algebra of statements}. Many of the methods utilized henceforth will be based on these tools.

\subsection{An ordering of statements}

Equivalence is one important relation between statements which we will utilize heavily. Next, we will define a partial ordering on the set of equivalence classes which will allow us to compare nonequivalent statements. While the whole set $\mathcal{S}$ is not partially ordered by narrowness, many of the important properties will still hold.

\begin{defn}\label{def_statement_narrowness_and_compatibility}
	Given two statement $\stmt_1$ and $\stmt_2$, we say that:
	\begin{itemize}
		\item $\stmt_1$ \emph{is narrower than} $\stmt_2$ (noted $\stmt_1 \narrower \stmt_2$) if $\stmt_1 \AND \NOT \stmt_2 \equiv \contradiction$.
		\item $\stmt_1$ \emph{is broader than} $\stmt_2$ (noted $\stmt_1 \broader \stmt_2$) if $\stmt_2 \narrower \stmt_1$.
		\item $\stmt_1$ \emph{is compatible to} $\stmt_2$ (noted $\stmt_1 \comp \stmt_2$) if $\stmt_1 \AND \stmt_2 \nequiv \contradiction$.

	\end{itemize}
	The negation of these properties will be noted by $\nnarrower$, $\nbroader$ , $\ncomp$ respectively.
\end{defn}

These relations capture a notion of conditional statement based on content rather than truth value. If $\stmt_1\narrower\stmt_2$, then $\stmt_2$ is true whenever $\stmt_1$ is true simply because of their content. If $\stmt_1\comp\stmt_2$, then their content allows them to be true at the same time (regardless of the actual truth value). A broader kind of compatibility will also be useful to define.

\begin{defn}\label{def_independent_statements}
	The elements of a set of statements $S \subseteq \mathcal{S}$ are said to be \textbf{independent} (noted $\stmt_1 \indep \stmt_2$ for a set of two) if any combination of their possibilities yields a consistent truth assignment. That is, $$\possFn(f(S)) = f_{\mathbb{B}}(\bigtimes\limits_{\stmt \in S} \possFn(\stmt))$$ for any Boolean function $f_{\mathbb{B}} : \mathbb{B}^{|S|} \to \mathbb{B}$ and associated function on statements $f:\mathcal{S}^{|S|}\to\mathcal{S}$. The negation of independence will be denoted by $\nindep$.
\end{defn}

\begin{prop}\label{prop_narrowness_properties}
	The above operations obey the following relationships:
	\begin{enumerate}
		\item 	$\stmt_1 \narrower \stmt_2$ if and only if $\stmt_1 \AND \stmt_2 \equiv \stmt_1$
		\item 	$\stmt_1 \narrower \stmt_2$ if and only if $\stmt_1 \OR \stmt_2 \equiv \stmt_2$
		\item 	$\stmt_1 \ncomp \stmt_2$ if and only if $\stmt_1 \AND \NOT \stmt_2 \equiv \stmt_1$
		\item 	$\stmt_1 \narrower \stmt_1 \OR \stmt_2$
		\item 	$\stmt_1 \AND \stmt_2 \narrower \stmt_1$
		\item $\stmt_1\narrower\stmt_2$ if and only if $\NOT\stmt_2\narrower\NOT\stmt_1$
		\item $\stmt_1\narrower\stmt_3$ and $\stmt_2\narrower\stmt_3$ implies that $\stmt_1\OR\stmt_2\narrower\stmt_3$
	\end{enumerate}	
\end{prop}

\begin{proof}
	For (i), consider $\stmt_1 \AND \stmt_2 \equiv \stmt_1 \AND \stmt_2 \OR \contradiction$. Since $\stmt_1 \narrower \stmt_2$, we have $\stmt_1 \AND \NOT \stmt_2 \equiv \contradiction$. Then $$\stmt_1 \AND \stmt_2 \OR \contradiction \equiv ( \stmt_1 \AND \stmt_2 ) \OR (\stmt_1 \AND \NOT \stmt_2) \equiv \stmt_1 \AND (\stmt_2 \OR \NOT \stmt_2) \equiv \stmt_1 \AND \tautology \equiv \stmt_1.$$ Therefore $\stmt_1 \AND \stmt_2 \equiv \stmt_1$. The same logic can be applied in reverse. The proof of (ii) is very similar.

	For (iii), consider $\stmt_1 \AND \NOT \stmt_2 \equiv \stmt_1 \AND \NOT \stmt_2 \OR \contradiction$. Since $\stmt_1 \ncomp \stmt_2$, we have $\stmt_1 \AND \stmt_2 \equiv \contradiction$. Then $$\stmt_1 \AND \NOT \stmt_2 \OR \contradiction \equiv ( \stmt_1 \AND \NOT \stmt_2 ) \OR (\stmt_1 \AND \stmt_2) \equiv \stmt_1 \AND (\NOT \stmt_2 \OR \stmt_2) \equiv \stmt_1 \AND \tautology \equiv \stmt_1.$$ Therefore $\stmt_1 \AND \NOT \stmt_2 \equiv \stmt_1$. The same logic can be applied in reverse.

	For (iv), we have $\stmt_1 \AND \NOT (\stmt_1 \OR \stmt_2) \equiv \stmt_1 \AND \NOT \stmt_1 \AND \NOT \stmt_2 \equiv \contradiction \AND \NOT \stmt_2 \equiv \contradiction$. Therefore $\stmt_1 \narrower \stmt_1 \OR \stmt_2$.

	For (v), we have $\stmt_1 \AND \stmt_2 \AND \NOT \stmt_1 \equiv \stmt_1 \AND \NOT \stmt_1 \AND \stmt_2 \equiv \contradiction \AND \stmt_2 \equiv \contradiction$. Therefore $\stmt_1 \AND \stmt_2 \narrower \stmt_1$.

	For (vi), if $\stmt_1\narrower\stmt_2$, then $\stmt_1\AND\NOT\stmt_2\equiv\contradiction$. But this is the same as $\NOT\stmt_2 \and \NOT(\NOT\stmt_1)\equiv\contradiction$, which is the same as $\NOT\stmt_2\narrower\NOT\stmt_1$. The reverse direction is similar.

	For (vii), note that by distributivity, $(\stmt_1\OR\stmt_2)\AND\NOT\stmt_3 \equiv (\stmt_1\AND\NOT\stmt_3)\OR(\stmt_2\AND\NOT\stmt_3) \equiv \contradiction \AND\contradiction\equiv \contradiction$.

\end{proof}

We are now ready to show several important properties of these relations, in particular that narrowness is a partial order on the set of equivalence classes.

\begin{prop}\label{prop_narrowness_is_if}
	Let $\stmt_1, \stmt_2 \in \mathcal{S}$ be two statements such that in all consistent truth assignments, if $\stmt_1$ is assigned $\TRUE$, then $\stmt_2$ must also be assigned $\TRUE$. Then $\stmt_1 \narrower \stmt_2$.
\end{prop}

\begin{proof}
	Consider a truth assignment $\{t_1, t_2, t_3\}$ for the sequence of statements $\{\stmt_1, \stmt_2, \stmt_1 \AND \NOT\stmt_2\}$. In order for it to be consistent, it must be either $\{\TRUE, \TRUE, \FALSE\}$, $\{\FALSE, \TRUE, \FALSE\}$ or $\{\FALSE, \FALSE, \FALSE\}$. As there is no assignment for $t_3=\TRUE$, we must have $\possFn(\stmt_1 \AND \NOT\stmt_2) = \{\FALSE\}$. Therefore $\stmt_1 \narrower \stmt_2$.
\end{proof}

\begin{prop}
	Statement narrowness satisfies the following properties:
	\begin{itemize}
		\item reflexivity: $s \narrower s'$ for all $s\equiv s'$
		\item antisymmetry: if $s_1 \narrower s_2$ and  $s_2 \narrower s_1$, then $s_1 \equiv s_2$
		\item transitivity: if $s_1 \narrower s_2$ and $s_2 \narrower s_3$, then $s_1 \narrower s_3$
	\end{itemize}
	and is therefore a \textbf{partial order}.
\end{prop}
\begin{proof}
Reflexivity is clear.

	For antisymmetry, suppose $\stmt_1 \narrower \stmt_2$ and $\stmt_1 \broader \stmt_2$. Then $\stmt_1 \AND \stmt_2 \equiv \stmt_1$ since $\stmt_1 \narrower \stmt_2$ and $\stmt_1 \AND \stmt_2 \equiv \stmt_2$ since $\stmt_2 \narrower \stmt_1$. Therefore $\stmt_1 \equiv \stmt_2$.

	For transitivity, let $\{t_1, t_2, t_3, t_4\}$ be a consistent truth assignment for the sequence of statements $\{\stmt_1, \stmt_2, \stmt_3, \stmt_1 \AND \NOT\stmt_3\}$.  Since $s_1 \narrower s_2$ and $s_2 \narrower s_3$, the consistent truth assignment must be one of the following: $\{\TRUE, \TRUE, \TRUE, \FALSE\}$, $\{\FALSE, \TRUE, \TRUE, \FALSE\}$, $\{\FALSE, \FALSE, \TRUE, \FALSE\}$, $\{\FALSE, \FALSE, \FALSE, \FALSE\}$.
	Since in all cases $t_4=\FALSE$ then $\possFn(\stmt_1 \AND \NOT\stmt_3) = \{\FALSE\}$. Therefore $\stmt_1 \narrower \stmt_3$.
\end{proof}

We can now exhibit some lattice-type properties of $\mathcal{S}$.

\begin{prop}
	Every subset $S \subseteq \stmtSet$ of statements has a supremum. That is, there exists an element $\bar{\stmt} \in \stmtSet$ such that $\stmt \narrower \bar{\stmt}$ for all $\stmt \in S$. 
\end{prop}

\begin{proof}
	Let $S \subseteq \stmtSet$ be an arbitrary set of statements. Using the Construction axiom, consider $\bar{\stmt}=\bigOR\limits_{\stmt[e] \in S} \stmt[e]$. Let $\stmt \in S$. Using the properties in \ref{boolean_properties}, one has $\stmt \OR \bar{\stmt} \equiv \bar{\stmt}$. By Propostion \ref{prop_narrowness_properties}, we have $\stmt \narrower \bar{\stmt}$ for any $\stmt \in S$.

Next, suppose $\stmt_0\in\mathcal{S}$ also has $\stmt\narrower\stmt_0$ for all $\stmt\in S$. Then by Proposition \ref{prop_narrowness_properties} (vii), it follows that $\bar{\stmt}\narrower\stmt_0$. Thus $\bar{\stmt}$ is our supremum. By antisymmetry of $\narrower$, it is unique up to statement equivalence. 
\end{proof}

\begin{cor}
	Every subset $S \subseteq \stmtSet$ of statements has an infimum. That is, there exists an element $\underline{\stmt} \in \stmtSet$ such that $\underline{\stmt} \narrower \stmt$ for all $\stmt \in S$. 
\end{cor}
\begin{proof}
Let $S\subseteq\mathcal{S}$ be a set of statements. consider the set $\NOT S$ consisting of all negations of elements of $S$. Then let $\stmt_0 = \sup\NOT S$. Then since $\NOT\stmt \narrower \stmt_0$ for all $\stmt\in S$, using Prop \ref{boolean_properties} we have that $\NOT\stmt_0\narrower\stmt$ for all $\stmt\in S$, and similar to the above proof, it is a greatest lower bound. We conclude that $\underline{\stmt} =\NOT\stmt_0$ is the infimum we sought. By antisymmetry of $\narrower$, it is unique up to statement equivalence. 
\end{proof}

It follows that the set of equivalence classes of statements is a \emph{complete} Boolean algebra. The supremum and infimum are constructed in the set of statements, but they are only unique up to statement equivalence.

We will now review the \emph{disjunctive normal form} which will give us some useful tools for analyzing large statements. While it is generally used in the context of more traditional logical systems requiring finite sentences, we will describe it in terms of arbitrary collections of statements. For a more extensive discussion and proofs of the finite case, see \S4.24 of \cite{lattices}.

\begin{defn}
\label{def_minterm}
Let $\{t_{\lambda}\}_{\lambda\in\Lambda}$ be an indexed collection of Booleans. A \emph{minterm} of $\{t_{\lambda}\}_{\lambda\in\Lambda}$ is a conjunction where each element appears only once, either negated or not. That is, it can be written as $m = \bigAND \limits_{\lambda\in\Lambda} (\NOT)^{a_{\lambda}} \, t_{\lambda}$ where $a_{\lambda} \in \mathbb{B}$, $\NOT ^ \TRUE \, t_{\lambda} := t_{\lambda}$ and $\NOT ^ \FALSE \, t_{\lambda} := \NOT t_{\lambda}$.
\end{defn}
We will often think of a minterm $m$ as a function of the Booleans, denoted $m(\{t_{\lambda}\}_{\lambda\in\Lambda})$. 
\begin{defn}
\label{def_dnf}
A Boolean function $f:\mathbb{B}^{\Lambda}\to\mathbb{B}$ is said to be in \emph{disjunctive normal form} (DNF) if it is expressed as a disjunction of minterms of the arguments, that is, if it is written as
 $$f(\{t_{\lambda}\}_{\lambda\in\Lambda}) =\bigOR \limits_{\nu\in X} \left( \bigAND \limits_{\lambda\in\Lambda} (\NOT)^{a_{\mu,\lambda}} \, t_{\lambda} \right)$$ where $X$ is another indexing set and  $a_{\mu,\lambda} \in \mathbb{B}$. 
\end{defn}

It is easy to see that a minterm $m$ will be true if and only if $t_{\lambda}=a_{\lambda}$ for all $\lambda\in\Lambda$. Note that in the definition of DNF, the cardinality of $X$ is exactly the cardinality of the set of inputs which yield $\TRUE$ as their output. We will use the same terminology \emph{minterm} and \emph{disjunctive normal form} for conjunctions of (possibly negated) distinct statements and disjunctions of minterms of statements, respectively. It is a useful fact that any Boolean function has a disjunctive normal form. This is classical for the finite case, and straightforward for the case of arbitrary size statements.

\begin{prop}\label{prop_dnf}
Any Boolean function $f:\mathbb{B}^{\Lambda}\to\mathbb{B}$ can be expressed in a DNF. 
\end{prop}
\begin{proof}
Each input of $f$ which yields $\TRUE$ corresponds to a minterm: if $f(\{a_{\lambda}\}_{\lambda\in\Lambda})=\TRUE$, then $$m(\{t_{\lambda}\}_{\lambda\in\Lambda}) = \AND_{\lambda\in\Lambda}(\NOT)^{a_{\lambda}}t_{\lambda}$$ is the associated minterm, since $f$ returns $\TRUE$ on the input $\{t_{\lambda}\}_{\lambda\in\Lambda}$ with $t_{\lambda}=a_{\lambda}$. Let $M$ be the set of minterms associated to outputs of $\TRUE$ for $f$. Then we have
$$
f(\{t_{\lambda}\}_{\lambda\in\Lambda}) = \bigOR\limits_{m\in M}m(\{t_{\lambda}\}_{\lambda\in\Lambda}).
$$
By construction, all $\TRUE$ outputs are achieved on the appropriate inputs since we include all the necessary minterms. No incorrect $\TRUE$ outputs are added, since each minterm will only return $\TRUE$ on its specific input. 
\end{proof}

By the Construction axiom, Proposition \ref{prop_dnf} allows us to express statement functions coming from Boolean functions in DNF as well.

\section{Verifiable statements and Experimental domains}
\label{verifiable}
Now that we have defined our basic structures of statements, we are ready to hone in on the specific types of statements which are most relevant to scientific inquiry, namely, those whose truth values can be discovered experimentally. The definition of an experimental test will not be made totally mathematically precise, but we will utilize the concept only to justify some of the definitions and axioms relating to verifiable statements, rather than as an actual tool in the proofs.

\begin{defn}\label{verifiable_statements}
The collection of \emph{verifiable statements} is a subset of the collection of statements, denoted by $\vstmtSet$. 
\end{defn}

The remark and axioms below will make it a bit more clear what kinds of statements belong in $\vstmtSet$.

\begin{rem}
The scientific application at hand will determine exactly which statements are verifiable, but a general way to think about which statements count as verifiable is as follows. An \emph{experimental test}, denoted by $\expt$, is a repeatable procedure which either terminates successfully, terminates unsuccessfully, or never terminates. We do not assume we can run arbitrarily many tests simultaneously. A statement $\stmt$ is said to be \emph{verifiable} if there exists an experimental test $\expt$ such that if $\truth(\stmt)=\TRUE$, then $\expt$ terminates successfully in finite time. This is not meant to be a precise definition, as it is too broad and does not quite capture the connection (which is more philosophical than mathematical) between a scientific test and a scientific theory. However, the general notion models an idealized version of the scientific method. Making this definition more rigorous is neither straightforward nor insightful for our present aims, so we will pursue it no further.

For the remainder of this section, we will include a series of remarks using the language of experimental tests to give non-rigorous justifications for some of the axioms. However, we will not rely on these for the actual proofs. 
\end{rem}

\begin{axiom}[Closure axioms]
\begin{enumerate}
\item	The conjunction of a finite collection of verifiable statements is a verifiable statement. 
\item The disjunction of a countable collection of verifiable statements is a verifiable statement. 
\end{enumerate}
\end{axiom}

\begin{rem}
The first of the Closure axioms is justified by the fact that if we can test each individual statement to be true, then we can simply run all of the (finitely many) tests to see if the larger statement is true. If any of the individual tests fail, or fail to terminate, then we will not verify the new statement. This cannot be generalized to infinite conjunctions, since there is no way to run infinitely many tests for an arbitrarily long time.

The second is somewhat more subtle. We can prescribe a recipe using a standard diagonal approach so that each of the countably many tests is eventually run for arbitrarily long. For example, we can run the first $n$ tests for $n$ seconds, starting with $n=1$ and incrementing, until some test terminates successfully. This will terminate successfully whenever any of the statements in the countable disjunction is true, which is sufficient to verify the new statement. It will only fail or fail to terminate if none of the statements' tests terminate successfully. This cannot be generalized beyond countable infinity since there is no procedure which will eventually run every test. 
\end{rem}

The collection $\vstmtSet$ of verifiable statements is not required to be closed under negation. To see why, consider for example the statement ``particles $A$ and $B$ are of different mass". If this statement is true, we could arrange for a sequence of progressively more precise measuring equipment to compare them, and eventually we would find out. Now, the negation of this statement, ``particles $A$ and $B$ have precisely the same mass" cannot be verified experimentally, since for any measuring apparatus, we will not know if the difference is just within a margin of error. However, we will be interested in certain subsets of $\vstmtSet$ with stronger closure properties.

\begin{defn}
The set of \emph{decidable statements}, denoted $\dstmtSet$, is the subset of $\vstmtSet$ consisting of statements $\stmt\in\vstmtSet$ such that $\NOT\stmt\in\vstmtSet$ as well. 
\end{defn}

\begin{rem}
Decidable statements may be thought of as those which can be shown to be true or false experimentally, or equivalently, as those pairs consisting of a statement and its negation which can both be verified. Given a pair of statements in $\vstmtSet$ which are negations of each other, we can simply run their associated tests for arbitrarily long (using e.g. a similar diagonalization approach to above) to see which one terminates successfully. 
\end{rem}

\begin{prop}\label{prop_decidable_AND_OR}
The collection $\dstmtSet$ of decidable statements is closed under finite conjunction and disjunction.
\end{prop}
\begin{proof}
	Let $\{\stmt_i\}_{i=1}^{n}$ be a finite collection of decidable statements. Then $$\{\stmt_i\}_{i=1}^{n} \subseteq \vstmtSet\text{ and }\{\NOT\stmt_i\}_{i=1}^{n} \subseteq \vstmtSet$$ are each collections of verifiable sets. Now, consider the conjunction $\bigAND\limits_{i=1}^{n} \stmt_i$. This a verifiable statement by the Closure axiom. Further, the negation $\NOT \bigAND\limits_{i=1}^{n} \stmt_i \equiv \bigOR\limits_{i=1}^{n} \NOT \stmt_i$ is also verifiable by the Closure axiom. We conclude that $\bigAND\limits_{i=1}^{n} \stmt_i\in\dstmtSet$.

	Similarly, the disjunction $\bigOR\limits_{i=1}^{n} \stmt_i$ and its negation $\NOT \bigOR\limits_{i=1}^{n} \stmt_i \equiv \bigAND\limits_{i=1}^{n} \NOT \stmt_i$ are verifiable statements by the closure axioms.
\end{proof}

Given the truth values of some collection of statements, we can immediately deduce the truth values of all statements constructed from them using conjunction and disjunction. We formalize this in the following definitions.

\begin{defn}
Given a set $\edomain\subseteq\vstmtSet$ of verifiable statements, $\basis\subseteq\edomain$ is a \emph{basis} for $\edomain$ if all elements of $\edomain$ can be generated from $\basis$ using finite conjunction and countable disjunction. 
\end{defn}

\begin{defn}
An \emph{experimental domain} $\edomain$ is a set of verifiable statements, closed under finite conjunction and countable disjunction, that includes precisely the tautology, the contradiction, that can be generated from a countable basis.
\end{defn}

Experimental domains represent all experimental knowledge that can be acquired about a scientific subject given unlimited time and resources. Once the truth values of the basis are known, then the entire collection is determined. The requirement of a countable basis comes from the fact that while we cannot test arbitrarily many statements simultaneously, given enough time, we can get to any individual test. In real-world applications, we can only ever perform finitely many tests, but in our idealized model we can always perform additional tests. Hence we allow for infinite bases, but only countably infinite.

Our eventual goal is to hone in on individual statements which determine all others in a domain. However, these will not necessarily be verifiable, and thus not within the actual experimental domain. We extend experimental domains to theoretical domains which will contain these special statements.

\begin{defn}
The \emph{theoretical domain} $\tdomain$ of an experimental domain $\edomain$ is the smallest set of statements containing $\edomain$ which is also closed under negation. Elements of a theoretical domain are called \emph{theoretical statements}.
\end{defn}

In particular, $\tdomain$ is the set of all statements generated from $\edomain$ using negation, countable conjunction, and countable disjunction. Because these are generated from $\basis$ by (at most) countable conjunction and disjunction, it follows that the truth values for all statements in a basis $\basis$ for an experimental domain $\edomain$ determine the truth values for all statements in $\tdomain$ as well as $\edomain$.

The theoretical domain can be thought of as the collection of statements (not necessarily verifiable) that can be used to state predictions. Certain kinds of statements with exact precision will fall in this domain, for example a claim such as ``particle $A$ has mass exactly 1g". There is an additional subset of the theoretical domain which will be useful when considering separability properties for topological spaces.

\begin{defn}\label{approx_verifiable}
A theoretical statement $\stmt\in\tdomain$ is \emph{approximately verifiable} if there exists a sequence $\{\obs_i\}_{i=1}^{\infty} \in \edomain$ such that $\stmt = \bigAND\limits_{i=1}^{\infty} \obs_i$.
\end{defn}

We may assume that $\stmt_{i+1}\narrower\stmt_i$ since if a statement is approximately verifiable, then in the notation above we can define $$\stmt_n' := \bigAND\limits_{j=1}^n\stmt_i.$$ By Corollary \ref{boolean_properties}, the conjunction of the $\stmt'_n$'s is equivalent to that of the original $\stmt_i$'s, and by Proposition \ref{prop_narrowness_properties} the $\stmt'_n$'s are increasingly narrow. Approximately verifiable statements give us a notion of a limit of statements, which we can think of in applications as a battery of increasingly precise experimental tests.

It is worth noting that the restriction to countable operations does leave out a large collection of statements. However, it would be impossible to even define a procedure to test a statement generated from uncountably many logic operations, so these should not be considered scientifically meaningful statements. Now that we have catalogued all the meaningful statements for a particular system, we are ready to define a very special collection of theoretical statements, which will be the last fundamentally new definition in this paper.

\section{Possibilities of an experimental domain}
\label{possibilities}

\begin{defn}
A \emph{possibility} for an experimental domain $\edomain$ is a theoretical statement $x\in\tdomain$ such that $x \nequiv \contradiction$ and for each $\mathsf{s} \in \tdomain$, either $x \narrower \mathsf{s}$ or $x \ncomp \mathsf{s}$. 
\end{defn}

We will generally denote possibilities by $x$, the set of all possibilities for an experimental domain by $X$. Since possibilities are statements, we use the same equivalence relation and ordering. The set of equivalence classes in $X$ will be denoted by $X^*$. We will often use the phrasing, ``a possibility $x$ \emph{determines} the truth value of a statement". This follows from the fact that if a possibility is true (or assigned true in a consistent assignment), then all statements incompatible must be assigned false, and all statements compatible are necessarily broader and hence must be assigned true. We now show a few of the fundamental properties of possibilities.

\begin{prop}\label{prop_poss_is_minterm}
	Let $\edomain$ be an experimental domain. A possibility for $\edomain$ is any minterm of a basis that is not a contradiction.
\end{prop}

\begin{proof}
	Let $\basis \subseteq \edomain$ be a basis for $\edomain$, and let $x$ be a minterm of $\basis$. Any theoretical statement $\stmt \in \tdomain$ can be expressed as the disjunction of minterms of $\basis$ by \ref{prop_dnf}. If $x$ is among the minterms needed to express $\stmt$, then $x \AND \stmt \equiv x$ and therefore $x \narrower \stmt$. If it is not among them, then $x \AND \NOT \stmt \equiv x$ and therefore $x \ncomp \stmt$ by Proposition \ref{prop_narrowness_properties}. Therefore a minterm is either narrower or incompatible with all theoretical statements and, if it is not a contradiction, it is a possibility by definition.

	Conversely, suppose $x \in \tdomain$ is a possibility. As it is a theoretical statement, it can be expressed as a disjunction of minterms of a basis $\basis$. Any minterm in such a disjunction would be narrower than $x$, but since $x$ is a possibility then it must be equivalent to $x$ by antisymmetry of narrowness. We conclude that $x$ must be expressed by a single minterm.
\end{proof}

\begin{prop} \label{no_other_poss}
Any statement $x\in\stmtSet$ other than a contradiction, such that for all $\stmt\in\tdomain$, $x\narrower\stmt$ or $x\ncomp\stmt$, and for which this fails for all nonequivalent statements $\stmt\notin\tdomain$, must be an element of the theoretical domain $\tdomain$. 
\end{prop}
\begin{proof}
A statement may be written as the conjunction of all statements that it determines (with appropriate negations). Determining the truth values of all statements in a theoretical domain is equivalent to determining the values on a (countable) basis $\basis$. As $x$ determines the truth values of all $\stmt\in\tdomain$, but no other statements, it may be written as a conjunction of the statements in $\basis$, with appropriate negations. But $\basis\subseteq\tdomain$ and $\tdomain$ is closed under countable conjunction and negation. 
\end{proof}

\begin{prop}\label{exactly_one_poss}
In the set of possibilities $X$ for any experimental domain $\edomain$, exactly one equivalence class of possibilities is true. 
\end{prop}
\begin{proof}
First, we show at most one possibility is true. Suppose $x_1,x_2\in X$ are both true. Since they are true simultaneously, they are not incompatible. Then we have $x_1\narrower x_2$ and $x_2\narrower x_1$, and so $x_1\equiv x_2$.

Next, we show at least one possibility is true. Let $X=\{x_{\lambda}\}_{\lambda\in\Lambda}$ be an indexed collection of all the possibilities, and suppose they are all false. Consider the new statement
$$
x' = \bigAND_{\lambda\in\Lambda}\NOT x_{\lambda}
$$
This is, by assumption, a true statement. By Proposition \ref{prop_poss_is_minterm} it is made from a conjunction of negations of minterms of the basis $\basis$ (in fact, this is in conjunctive normal form). Because it is true, there must be a consistent truth assignment for some of the elements of $\basis$ such at (at least) one term in each disjunction $\NOT x_{\lambda}$ is true. By the Completeness axiom, this may be extended to a consistent truth assignment of all statements in $\basis$. Let $y$ be the conjunction of the elements of $\basis$ with the appropriate negations to match this consistent truth assignment. Then this is a non-contradictory (since it is a conjunction made from a consistent truth assignment) minterm of the basis elements, and it is distinct from all other possibilities. This contradicts the assumption that $X$ contained all possibilities.
\end{proof}

Proposition \ref{prop_poss_is_minterm} allows us to explicitly describe the possibilities of any domain in terms of a basis. Note that not all possibilities are approximately verifiable (a countable conjunction of verifiable statements), because possibilities may include negations of basis elements, and not all negations are still verifiable. Proposition \ref{no_other_poss} essentially tells us that the requirement that possibilities be theoretical statements is vacuous. However, it is very useful to keep track of the theoretical domain, as it gives us a concrete space from which the possibilities are to be drawn, and it will serve other purposes later.

Finally, Proposition \ref{exactly_one_poss} motivates the study (and name) of possibilities: it shows us that possibilities are the most precise information one can define about a system under scientific study, and cover all possible outcomes of experiments. One special case is worth noting in particular: we have not excluded the possibility that no test will succeed.

	\begin{defn}
		The \emph{residual possibility} $\mathring{x}$ for an experimental domain $\edomain$ with basis $\basis$ is $$\mathring{x} = \bigAND\limits_{\stmt[e] \in \basis} \NOT e$$ if it is not a contradiction. An experimental domain is \textbf{complete} if it doesn't admit a residual possibility.
	\end{defn}

	\begin{defn}
	The \textbf{established possibilities} of a domain is the set $\dot{X}=X\setminus\{\mathring{x}\}$ of all possibilities excluding the residual possibility.
\end{defn}

The residual possibility represents the case that the methods available are insufficient. This possibility corresponds to what remains after all verifiable statements are tested unsuccessfully. Not all domains admit a residual possibilities: a domain is complete if there are sufficient verifiable statements to cover all outcomes. A simple example might be in counting a fixed finite collection of objects. An example of an incomplete domain might be in identifying organisms, since we have not catalogued all species. The established possibilities in that case would be a list of all known species, and the residual possibility would simply be that it is a new discovery.

With the tools we have thus far developed, we are now ready to answer a fundamental question: what is the size of the largest set amongst which scientific methods can distinguish? This is answered by looking at the possible cardinalities of sets of (equivalence classes of) possibilities. The requirement of a countable basis for an experimental domain is crucial. Recall Theorem \ref{cardinality_thm}, which states that $X^*$ has cardinality at most that of the continuum.

\begin{proof}[Proof of Theorem \ref{cardinality_thm}]
Because each possibility is equivalent to a non-contradictory minterm of the basis $\basis$, we can find an upper bound using the cardinality of the set of all minterms for $\basis$. Each minterm is determined precisely by a choice of $\TRUE$ or $\FALSE$ for each element of $\basis$. But $\basis$ is countable, so the set of minterms has cardinality $|2^{\N}|$, as desired.
\end{proof}

We have now developed all of the major objects for our theory, and are ready to put them together into a more robust mathematical structure.

\section{A topology for experimental domains}

Topology is a natural setting for axiomatizing scientific study since a motivating goal of topology is to formalize ``nearness" for points in a set without a precise metric. All experiments have limited precision, and so each test will single out a neighborhood rather than a point. We will use the tools from the previous sections to build a natural topology associated to an experimental domain with the possibilities as the set of points, and verifiable sets will correspond to neighborhoods. We will first associate natural subsets of the set of possibilities $X$ to verifiable sets in an experimental domain $\edomain$ with basis $\basis$.

\begin{defn}
	Let $\edomain$ be an experimental domain and $X$ its possibilities. Define the map $U : \edomain \rightarrow 2^X$ that for each statement $\obs \in \edomain$ returns the set of possibilities compatible with it: $$U(\obs)\equiv\{ x \in X \, | \, x \comp \obs\}.$$ We call $U(\obs)$ the \textbf{verifiable set} of possibilities associated with $\obs$.
\end{defn}

\begin{prop}
	A statement $\obs \in \edomain$ is equivalent to the disjunction of the possibilities in its verifiable set $U(\obs)$. That is, $\obs \equiv \bigOR\limits_{x \in U(\obs)} x$.
\end{prop}
\begin{proof}
Since $\stmt$ is a member of an experimental domain, it may be expressed in DNF as a disjunction of minterms of the basis $\basis$, and in particular, minterms which are not contradictions. This means $\stmt$ is a disjunction of possibilities, say $\stmt\equiv\bigOR\limits_{x \in U} x$. The statement $\stmt$ is broader than precisely the possibilities in its disjunction, and incompatible with all other possibilities. This means $U=U(\stmt)$, as desired. 
\end{proof}.

We are now ready to define the topology on $X^*$, the set of equivalence classes of possibilities.

\begin{prop}[Theorem \ref{main}, the topology]
\label{topology}
The collection of verifiable sets $\mathsf{T}_X=\{U(\stmt)|\stmt\in\edomain\}$ for $\stmt\in\edomain$ forms a topology on the set $X^*$. 
\end{prop}
\begin{proof}
First, recall that $\tautology$ and $\contradiction$ are verifiable sets. We have $U(\tautology) = X$ and $U(\contradiction) = \emptyset$.

Next, let $U(\stmt_1)$ and $U(\stmt_2)$ be verifiable sets for statements $\stmt_1$ and $\stmt_2$. Then $$U(\obs_1\AND\obs_2) = \{ x \in X \, | \, x \comp (\obs_1\AND\obs_2)\} =  \{ x \in X \, | \, x \comp \obs_1\} \cap \{ x \in X \, | \, x \comp \obs_2\} = U(\obs_1) \cap U(\obs_2)$$, and hence $\mathsf{T}_X$ is closed under pairwise (and hence finite) intersections.

Finally, let $\{U(\stmt_{\lambda})\}_{\lambda\in\Lambda}$ be an arbitrary collection of verifiable sets with each $\stmt_{\lambda}\in\edomain$. By Lemma \ref{countable_disjunction} (below), it will suffice to assume it is a countable collection. Then $$U(\bigOR\limits_{\lambda\in\Lambda}\stmt_{\lambda}) = \{ x \in X \, | \, x \comp (\bigOR\limits_{\lambda\in\Lambda}\stmt_{\lambda})\} =\bigcup\limits_{\lambda\in\Lambda} \{ x \in X \, | \, x \comp \stmt_{\lambda}\} = \bigcup\limits_{\lambda\in\Lambda} U(\obs_{\lambda}).$$ Because $\Lambda$ is countable, $\bigOR\limits_{\lambda\in\Lambda}\stmt_{\lambda}$ is a countable disjunction of verifiable sets, so it is still a verifiable set, and the left-hand side is a verifiable set for an element of $\edomain$.

Thus $T_X$ is closed under arbitrary unions, and so we conclude that it is a topology.
\end{proof}

\begin{lem}\label{countable_disjunction}
An arbitrary disjunction of verifiable statements in an experimental domain $\edomain$ may be re-expressed as a countable disjunction of statements in the basis $\basis$.
\end{lem}
\begin{proof}
Let $\bigOR\limits_{\lambda\in\Lambda}\stmt_{\lambda}$ be an arbitrary disjunction of verifiable statements in an experimental domain. Since all statements in $\edomain$ are generated by elements of the countable set $\basis$, we claim that $$\bigOR\limits_{\lambda\in\Lambda}\stmt_{\lambda} \equiv \bigOR\limits_{i=1}^{\infty}b_i$$ for some (countable) collection $\{b_i\}_{i=1}^{\infty}$ of verifiable statements which are finite conjunctions of basis elements.

To see why, notice that each $\stmt_{\lambda}$ is itself made from finite conjunction and countable disjunction of basis statements, so we can rearrange it into a DNF where each term is a finite conjunction. Then, the disjunction of all the $\stmt_{\lambda}$'s is a (possibly uncountable) disjunction of finite conjunctions of basis elements. But since there are only countably many basis elements to begin with, there can only be countably many distinct finite conjunctions of basis elements, which we label as $\{b_i\}_{i=1}^{\infty}$. 
\end{proof}

\begin{defn}
The topology $T_X$ defined in Proposition \ref{topology} is called the \emph{topology of experimental possibility}.
\end{defn}

The topology of experimental possibility classifies near-ness in all scientific scenarios. It will organize all scientific statements in a single domain be they taxonomic classifications or possible masses for an unknown object. It is a general construction for any area of scientific study. We can now study properties of this topology to better understand the kinds of possibilities that science can hope to distinguish. From the proof, it is clear that conjunction of verifiable statements correspond to intersection of verifiable sets, and likewise disjunction corresponds to union.

\begin{cor}[Theorem \ref{main}, countability]
The collection $U(\basis)\cup\{X\}$ of verifiable sets for a basis of an experimental domain, together with the set of all possibilities, forms a sub-basis for the topology of experimental possibility $T_X$.
\end{cor}
\begin{proof}
It is clear that $U(\basis)$ can generate all nontrivial open sets by union (countable, and hence arbitrary, disjunction) and finite intersection (conjunction). Because $\tautology$ is not necessarily in $\basis$, we need to add $\{X^*\}$ to the collection to reach the residual possibility, if it exists. 
\end{proof}

Note that if the domain $\edomain$ is complete, then $U(\basis)$ will be sufficient.

\begin{cor}[Theorem \ref{main}, countability]
The topology of experimental possibility is second-countable.
\end{cor}
\begin{proof}
The collection $U(\basis)\cup\{X^*\}$ is a countable sub-basis which may be extended to a countable basis by including finite intersections. 
\end{proof}

Next, we will explore separability properties. Recall that a topological space has $\mathsf{T}_0$ separability if no two points have exactly the same set of neighbors, and $\mathsf{T}_2$ (Hausdorff) separability if any two distinct points can be separated by open neighborhoods. Separability is a highly desirable property in scientific applications. If two possibilities share all the same neighborhoods, then no experimental test can distinguish them. This opens up a question of whether they can even be considered to be different. Fortunately, this is never the case in our theory.

\begin{prop}[Theorem \ref{main}, $\mathsf{T}_0$ separability]
The topology of experimental possibility has $\mathsf{T}_0$ separability for any experimental domain. 
\end{prop}
\begin{proof}
Let $x_1,x_2\in X^*$. It will suffice to show that there exists a verifiable statement $\mathsf{e}\in\basis$ whose verifiable set contains one but not the other. Because possibilities may be expressed as minterms of basis elements, if $x_1$ and $x_2$ are distinct, then there exists a basis element $b\in\basis$ which is negated in, without loss of generality, $x_1$ but not $x_2$. Then $x_2\in U(b)$ but $x_1\notin U(b)$, as required. 
\end{proof}

\begin{prop}[Theorem \ref{main}, $\mathsf{T}_2$ separability]
The topology of experimental possibility is Hausdorff (or $\mathsf{T}_2$) if and only if all possibilities are approximately verifiable.
\end{prop}
\begin{proof}
Suppose first that all possibilities are approximately verifiable. Let $$x_1=\bigAND\limits_{i=1}^\infty \obs_i^1\text{ and }x_2=\bigAND\limits_{j=1}^\infty \obs_j^2$$ be two nonequivalent possibilities, written as countable conjunctions of verifiable statements using approximate verifiability. We may assume the sequences in the conjunctions have the property that $\stmt^j_{i+1}\narrower\stmt^j_i$ for $j=1,2$ and all $i\in\N$.

We have $x_1\AND x_2\equiv\contradiction$. It follows that there must exist $m,n$ such that $\stmt^1_m\AND\stmt^2_n\equiv\contradiction$, since otherwise there would be a consistent truth assignment to all the $\stmt^k_l$'s allowing $x_1\AND x_2$ to be true. This means that $U(\stmt^1_m)\cap U(\stmt^2_n)=\emptyset$, and these are open sets containing $\stmt_1$ and $\stmt_2$ respectively, as required.

Next, suppose $T_X$ is Hausdorff. Let $x\in X$ be a possibility. Consider the (countable) collection of all verifiable sets in the basis containing $x$, denoted by $\{V_i\}_{i=1}^{\infty}$. The intersection $\cap_iV_i$ must be the singleton $\{x\}$ because the topology is Hausdorff. But the basis for the topology comes from the basis $\basis$ of the experimental domain $\edomain$. In particular, for each $i$ there exists $\stmt_i\in\basis\subseteq\edomain$ with $V_i=U(\stmt_i)$. It follows that $x=\bigAND\limits_{i=1}^{\infty}\stmt_i$, so $x$ is approximately verifiable. 
\end{proof}

Many experimental domains in application have Hausdorff topologies. All finite domains, as well as most reasonable domains based on numerical values will have approximately verifiable possibilities.

\begin{rem}
One can also define ``theoretical sets" of possibilities associated to theoretical statements with the same compatibility criterion. Using similar proof techniques, this gives rise to a $\sigma$-algebra. One can show, using the fact that negation corresponds to set-complement, that this $\sigma$-algebra the Borel $\sigma$-algebra for the topology of experimental possibility. We have chosen not to expand on this here, as the major applications of this $\sigma$-algebra are less developed and a bit beyond the goals of this paper. 
\end{rem}

These mathematical tools are well-justified for use in organizing scientific claims and experimental results. Any area of science can use these same ideas to characterize the relevant experimental domains. We have also provided a precise scientific interpretation to the mathematical ideas presented here.

\section{Conclusions and questions}
\label{conclusion}

We have seen the basic foundation of a mathematical theory for experimental distinguishability. This opens far more questions than it answers. For example, how different experimental domains can relate to each other can lead to a mathematical definition of causality. Defining discrete and continuous properties of objects can lead to special kinds of experimental domain with more structure. It seems within reach to recover the real numbers and the Euclidean topology (for experimental domains relating to measuring sizes, masses, etc.) using methods similer to those in this paper. A longer-term goal is to recover the structure of manifolds for physics in three-dimensional space. The evolving manuscript \cite{carc3} contains substantially more content than can fit in a single journal article, and includes progress on several of these next steps, in addition to numerous examples and discussions on the topics covered here.

This avenue of study seems largely unexplored. We hope to bring to light and possibly answer some of these fundamental questions concerning the nature of science, and the unreasonable effectiveness of mathematics in the sciences \cite{wigner}.

\bibliography{bibliography}

\begin{thebibliography}{1}
	\bibitem{carc1} G. Carcassi, C. A. Aidala, D. J. Baker and L. Bieri: From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics. Journal of Physics Communications, 2, 4, 045026, 2018.

	\bibitem{carc2} C. A. Aidala, G. Carcassi, M. J. Greenfield: \emph{Topology and experimental distinguishability}. arXiv:1708.05492, 2017.

	\bibitem{carc3} G. Carcassi, C. A. Aidala: \emph{Assumptions of Physics} (in preparation). http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.

	\bibitem{lattices} B. A. Davey, H. A. Priestley: \emph{Introduction to lattices and order}. Cambridge university press, 2002.

	\bibitem{wigner} E. P. Wigner: The unreasonable effectiveness of mathematics in the natural sciences. Mathematics and Science, pp. 291-306, 1990.

\end{thebibliography}

\end{document}

### 2018-10-02 Articles/LogicalTopologicalFramework/article.tex
Various changes made as we discussed

@@
 
-\begin{defn}
-A \emph{statement} is an object with a truth value, ascertained by a truth function outputting a Boolean, and logical content, defined by satisfying the Consistency Axiom (\ref{consistent}) and the Completeness Axiom (\ref{completeness}).
+We begin the axiomatization of our universe of scientific statements by declaring existence of our collection. Following this, the special properties of its elements are built up.
+
+\begin{axiom}[Existence Axiom]
+There exists a collection $\mathcal{S}$ of \emph{statements} whose properties are as follows.
+Each $\stmt\in\mathcal{S}$ is an object with a truth value, ascertained by a truth function outputting a Boolean, and logical content, defined by satisfying the Consistency Axiom (\ref{consistent}) and the Completeness Axiom (\ref{completeness}).
 The set of all statements is denoted $\mathcal{S}$.
 The truth function is denoted $\truth:\mathcal{S}\to\mathbb{B}$.
-\end{defn}
+\end{axiom}
 
@@
 Our earlier example of two statements about electron mass cannot both be true, but all other combinations (both false, one true and one false) are consistent truth assignments.
+We now can define the Completeness Axiom:
+
+\begin{axiom}[Completeness Axiom]
+\label{completeness}
+Given a collection of statements $\{\stmt_{\lambda}\}_{\lambda\in\Lambda}$ for some (possibly uncountable) index set $\Lambda$, let $A\subseteq\Lambda$, and let $\{t_a\}_{a\in A}$ be a consistent truth assignment for $\{\stmt_a\}_{a\in A}$.
+Then there exists a consistent truth assignment $\{t_{\lambda}\}_{\lambda\in\Lambda}$ for $\{\stmt_{\lambda}\}_{\lambda\in\Lambda}$ such that $t_{\lambda} = t_a$ for all $\lambda\in A$.
+\end{axiom}
+This axiom restricts the notion of what is considered to be a consistent truth assignment.
+No consistent truth assignment for any single statement can lead to a paradox in larger collections of statements.
 We next define the possibilities of a statement in terms of consistent truth values of single statements.
@@
 $$
-where $\possFn(\stmt)$ is defined to be the set $S\subseteq\mathbb{B}$ of truth values such that $\{\stmt\},\{B\}$ is a consistent truth assignment for each $B\in S$. 
+where $\possFn(\stmt)$ is defined to be the set $B\subseteq\mathbb{B}$ of truth values such that $\{\stmt\},\{t\}$ is a consistent truth assignment for each $t\in B$. 
 \end{defn}
@@
 The notion of possibilities of a statement captures the difference between statements whose truth values can be obtained a priori (such as mathematical statements or tautologies), and those which must be experimentally tested.
-It is immediate that $\truth(\stmt)\in\possFn(\stmt)$ for all statements $\stmt$.
-In fact, for any collection of statements, assigning them their truth values is always a consistent truth assignment, since we assume true scientific statements to be non-contradictory.
+For any collection of statements, assigning them their truth values is always a consistent truth assignment, since we assume true scientific statements to be non-contradictory.
 However, there are many consistent truth assignments which do not arise from the truth values of the included statements.
 Conversely, for any consistent truth assignment, each truth value $t_{\lambda}$ must be in $\possFn(\stmt_{\lambda})$.
-We now can define the Completeness Axiom:
 
-\begin{axiom}[Completeness Axiom]
-\label{completeness}
-Given a collection of statements $\{\stmt_{\lambda}\}_{\lambda\in\Lambda}$ for some (possibly uncountable) index set $\Lambda$, let $A\subseteq\Lambda$, and let $\{t_a\}_{a\in A}$ be a consistent truth assignment for $\{\stmt_a\}_{a\in A}$.
-Then there exists a consistent truth assignment $\{t_{\lambda}\}_{\lambda\in\Lambda}$ for $\{\stmt_{\lambda}\}_{\lambda\in\Lambda}$ such that $t_{\lambda} = t_a$ for all $\lambda\in A$.
-\end{axiom}
-This axiom expands the notion of what is considered to be a possible truth value.
-No possible truth value of any statement can lead to a paradox in larger collections of statements.
-These two axioms define our basic collection of statements.
+It is important to note that the possibilities of a statement composed from other statements cannot be determined just by looking at the truth values and possibilities of individual statements alone.
+For example, two statements $\stmt_1$ and $\stmt_2$ which both have $\mathbb{B}$ as their possibilities may have any subset of $\mathbb{B}$ as the possibilities of their conjunction $\stmt_1\AND\stmt_2$ (regardless of the truth values).
 Next, we will give an additional axiom which will allow us to create new statements from old ones.
@@
 \begin{axiom}[Construction axiom]
-Let $f_{\mathbb{B}}$ be a Boolean function $f_{\mathbb{B}}:\mathbb{B}^{\Lambda}\to\mathbb{B}$, where $\Lambda$ is any cardinal.
-Then there exists a function on the set of statements $f:(\mathcal{S}^*)^{\Lambda}\to\mathcal{S}$ such that
-		$$\truth(f(\stmt_1, ..., \stmt_n)) = f_{\mathbb{B}}(\truth(\stmt_1), ..., \truth(\stmt_n))$$
-with the same relationship holding for every consistent truth assignment. 
+Let $f_{\mathbb{B}}$ be a Boolean function $f_{\mathbb{B}}:\mathbb{B}^{\Lambda}\to\mathbb{B}$, where $\Lambda$ is an index set.
+Then there exists a function on the set of statements $f:\mathcal{S}^{\Lambda}\to\mathcal{S}$ with the following requirements.
+First, the truth value is determined by the Boolean function:
+\begin{equation}
+\label{constructionrequirement}
+\truth(f(\{\stmt_{\lambda}\}_{\lambda\in\Lambda}) = f_{\mathbb{B}}(\{\truth(\stmt_{\lambda})\}_{\lambda\in\Lambda}).
+\end{equation}
+Second, any consistent truth assignment for a collection of statements $S$ which includes the statement $f(\{\stmt_{\lambda}\}_{\lambda\in\Lambda})$ extends to a consistent truth assignment for $S\cup\{\stmt_{\lambda}\}_{\lambda\in\Lambda}$ such that the assignment to $f(\{\stmt_{\lambda}\}_{\lambda\in\Lambda})$ is given by $f_{\mathbb{B}}(\{t_{\lambda}\}_{\lambda\in\Lambda}$.
 \end{axiom}
 
-The point of the Construction axiom is to extend our usual Boolean logical operations to statements. We will now define them in this context.
+\begin{rem}
+Note that in the second condition of the axiom, this means that if $S$ already contains some of the $\stmt_{\lambda}$'s, then their truth assignments must already be compatible in the sense of the second condition above.
+\end{rem}
+
+The Construction axiom allows us to turn any Boolean function into a function on the collection of statements. In particular, the usual logic operations can be applied to $\mathcal{S}$.
 
@@
 
-It is important to note that the possibilities of a statement composed from other statements cannot be determined just by looking at the truth values and possibilities of individual statements alone.
-For example, two statements $\stmt_1$ and $\stmt_2$ which both have $\mathbb{B}$ as their possibilities may have any subset of $\mathbb{B}$ as the possibilities of their conjunction $\stmt_1\AND\stmt_2$ (regardless of the truth values).
-
 Now, we will define an equivalence relation on the set of statements which will dramatically reduce the collection of statements to a more manageable size.
@@
 \begin{defn}
-Two statements $\stmt_1$ and $\stmt_2$ are \emph{equivalent}, denoted $\stmt_1 \equiv \stmt_2$, if and only if $$\stmt_0:=(\stmt_1 \AND \stmt_2) \OR (\NOT\stmt_1 \AND \NOT\stmt_2)$$ has $\possFn(\stmt_0)=\{\TRUE\}$.
+Two statements $\stmt_1$ and $\stmt_2$ are \emph{equivalent}, denoted $\stmt_1 \equiv \stmt_2$, if and only if in all consistent truth assignments, $\stmt_1$ and $\stmt_2$ must be assigned the same truth value.
 \end{defn}
@@
 
+It is an immediate consequence of the definition that all tautologies are equivalent, and all contradictions are equivalent.
+We will interchangeably denote tautologies and their equivalence class by $\tautology$, and similarly contradictions and their equivalence class will be denoted by $\contradiction$.
 An alternative way to view statement equivalence is as follows:
@@
 
-It is an immediate consequence of the definition that all tautologies are equivalent, and all contradictions are equivalent.
-We will interchangeably denote tautologies and their equivalence class by $\tautology$, and similarly contradictions and their equivalence class will be denoted by $\contradiction$.
 Proposition \ref{prop_equivalent_is_iff} allows one to quickly see that if two Boolean functions are equal, then the corresponding functions on $\mathcal{S}$ must give equivalent results.
@@
 
-While $\mathcal{S}$ is not technically a Boolean algebra itself (since e.g.
-there are many nonequal, but equivalent, tautologies and contradictions), it will enjoy many of the useful consequences of being a Boolean algebra when using statement equivalence rather than equality.
-We call the operations and properties defined thus far the \emph{algebra of statements}.
-Many of the methods utilized henceforth will be based on these tools.
+Note that $\mathcal{S}$ is not a Boolean algebra itself (since e.g.
+there are many nonequal, but equivalent, tautologies and contradictions).
+Henceforth, unless otherwise noted, \emph{we will exclusively use the notation $\mathcal{S}$ for the collection of equivalence classes of statements}.
+The Boolean operations with $\tautology$ and $\contradiction$ (now each individual equivalence classes) endow the set of equivalence classes $\mathcal{S}$ with the structure of a Boolean algebra.
+Further, it is a \emph{complete Boolean algebra}, since the usual operations of conjunction and disjunction allow one to obtain an $\inf$ and $\sup$ for any subset $S\subseteq\mathcal{S}$.
+
+We call the operations and properties defined thus far the \emph{algebra of statements}. Many of the methods utilized henceforth will be based on these tools.
 
@@
 
-We can now exhibit some lattice-type properties of $\mathcal{S}$.
+We can now exhibit some lattice-type properties of $\mathcal{S}$ endowed with $\narrower$.
 
 \begin{prop}
-	Every subset $S \subseteq \stmtSet$ of statements has a supremum. That is, there exists an element $\bar{\stmt} \in \stmtSet$ such that $\stmt \narrower \bar{\stmt}$ for all $\stmt \in S$. 
+The supremum (respectively infimum) of any set $S \subseteq \stmtSet$ of statements with respect to $\narrower$ is equal to the Boolean supremum (respectively infimum).
+That is, for any collection $S$, the conjunction $\OR S$ (respectively $\AND S$ is the supremum (respectively infimum) with respect to $\narrower$.
+Denote them by $\OR S := \bar{\stmt} \in \stmtSet$ and $\AND S := \underline{\stmt}\in\stmtSet$.
 \end{prop}
@@
 By antisymmetry of $\narrower$, it is unique up to statement equivalence.
-\end{proof}
 
-\begin{cor}
-	Every subset $S \subseteq \stmtSet$ of statements has an infimum.
-That is, there exists an element $\underline{\stmt} \in \stmtSet$ such that $\underline{\stmt} \narrower \stmt$ for all $\stmt \in S$.
-\end{cor}
-\begin{proof}
-Let $S\subseteq\mathcal{S}$ be a set of statements.
-consider the set $\NOT S$ consisting of all negations of elements of $S$.
+For the infimum, consider the set $\NOT S$ consisting of all negations of elements of $S$.
 Then let $\stmt_0 = \sup\NOT S$.
@@
 \end{proof}
-
-It follows that the set of equivalence classes of statements is a \emph{complete} Boolean algebra.
-The supremum and infimum are constructed in the set of statements, but they are only unique up to statement equivalence.
 
@@
 Now, the negation of this statement, ``particles $A$ and $B$ have precisely the same mass" cannot be verified experimentally, since for any measuring apparatus, we will not know if the difference is just within a margin of error.
-However, we will be interested in certain subsets of $\vstmtSet$ with stronger closure properties.
-
-\begin{defn}
-The set of \emph{decidable statements}, denoted $\dstmtSet$, is the subset of $\vstmtSet$ consisting of statements $\stmt\in\vstmtSet$ such that $\NOT\stmt\in\vstmtSet$ as well. 
-\end{defn}
-
-\begin{rem}
-Decidable statements may be thought of as those which can be shown to be true or false experimentally, or equivalently, as those pairs consisting of a statement and its negation which can both be verified.
-Given a pair of statements in $\vstmtSet$ which are negations of each other, we can simply run their associated tests for arbitrarily long (using e.g.
-a similar diagonalization approach to above) to see which one terminates successfully.
-\end{rem}
-
-\begin{prop}\label{prop_decidable_AND_OR}
-The collection $\dstmtSet$ of decidable statements is closed under finite conjunction and disjunction.
-\end{prop}
-\begin{proof}
-	Let $\{\stmt_i\}_{i=1}^{n}$ be a finite collection of decidable statements.
-Then $$\{\stmt_i\}_{i=1}^{n} \subseteq \vstmtSet\text{ and }\{\NOT\stmt_i\}_{i=1}^{n} \subseteq \vstmtSet$$ are each collections of verifiable sets.
-Now, consider the conjunction $\bigAND\limits_{i=1}^{n} \stmt_i$.
-This a verifiable statement by the Closure axiom.
-Further, the negation $\NOT \bigAND\limits_{i=1}^{n} \stmt_i \equiv \bigOR\limits_{i=1}^{n} \NOT \stmt_i$ is also verifiable by the Closure axiom.
-We conclude that $\bigAND\limits_{i=1}^{n} \stmt_i\in\dstmtSet$.
-
-	Similarly, the disjunction $\bigOR\limits_{i=1}^{n} \stmt_i$ and its negation $\NOT \bigOR\limits_{i=1}^{n} \stmt_i \equiv \bigAND\limits_{i=1}^{n} \NOT \stmt_i$ are verifiable statements by the closure axioms.
-\end{proof}
 
@@
 We will generally denote possibilities by $x$, the set of all possibilities for an experimental domain by $X$.
-Since possibilities are statements, we use the same equivalence relation and ordering.
-The set of equivalence classes in $X$ will be denoted by $X^*$.
+Since possibilities are statements, we can use the same ordering.
 We will often use the phrasing, ``a possibility $x$ \emph{determines} the truth value of a statement".
@@
 \begin{prop}\label{exactly_one_poss}
-In the set of possibilities $X$ for any experimental domain $\edomain$, exactly one equivalence class of possibilities is true. 
+In the set of possibilities $X$ for any experimental domain $\edomain$, exactly one of the possibilities is true. 
 \end{prop}
@@
 With the tools we have thus far developed, we are now ready to answer a fundamental question: what is the size of the largest set amongst which scientific methods can distinguish?
-This is answered by looking at the possible cardinalities of sets of (equivalence classes of) possibilities.
+This is answered by looking at the possible cardinalities of sets of possibilities.
 The requirement of a countable basis for an experimental domain is crucial.
-Recall Theorem \ref{cardinality_thm}, which states that $X^*$ has cardinality at most that of the continuum.
+Recall Theorem \ref{cardinality_thm}, which states that $X$ has cardinality at most that of the continuum.
 
@@
 
-We are now ready to define the topology on $X^*$, the set of equivalence classes of possibilities.
+We are now ready to define the topology on $X$, the set of possibilities.
 
@@
 \label{topology}
-The collection of verifiable sets $\mathsf{T}_X=\{U(\stmt)|\stmt\in\edomain\}$ for $\stmt\in\edomain$ forms a topology on the set $X^*$. 
+The collection of verifiable sets $\mathsf{T}_X=\{U(\stmt)|\stmt\in\edomain\}$ for $\stmt\in\edomain$ forms a topology on the set $X$. 
 \end{prop}
@@
 It is clear that $U(\basis)$ can generate all nontrivial open sets by union (countable, and hence arbitrary, disjunction) and finite intersection (conjunction).
-Because $\tautology$ is not necessarily in $\basis$, we need to add $\{X^*\}$ to the collection to reach the residual possibility, if it exists.
+Because $\tautology$ is not necessarily in $\basis$, we need to add $\{X\}$ to the collection to reach the residual possibility, if it exists.
 \end{proof}
@@
 \begin{proof}
-The collection $U(\basis)\cup\{X^*\}$ is a countable sub-basis which may be extended to a countable basis by including finite intersections. 
+The collection $U(\basis)\cup\{X\}$ is a countable sub-basis which may be extended to a countable basis by including finite intersections. 
 \end{proof}
@@
 \begin{proof}
-Let $x_1,x_2\in X^*$.
+Let $x_1,x_2\in X$.
 It will suffice to show that there exists a verifiable statement $\mathsf{e}\in\basis$ whose verifiable set contains one but not the other.

### 2018-10-22 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Good first draft

--- 2017-07-14 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
@@
-\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
+\documentclass[11pt]{article}
+
+\usepackage[margin=1.25in]{geometry}
 
@@
 
-\title{DRAFT \\ Hamiltonian mechanics is conservation of information \\ (or why isolated systems conserve energy)}
+\title{Hamiltonian mechanics is conservation of information entropy}
 \author{Christine A. Aidala, Gabriele Carcassi}
@@
 
+\maketitle
+
 \begin{abstract}
@@
 
-\maketitle
+\tableofcontents
+\newpage
 
@@
 
-\section{Single degree of freedom}
-Define distribution over phase space.
+\section{Distributions}
 
-Show change under coordinate transformation.
+The general idea is that we want to describe a quantity that is distributed among different cases.
+For example, mass or charge over different positions in space, amount of material among different compounds, population at different ages.
+In all these cases, we assume there is a total quantity $M$, a set of $n$ variables $\xi^a$ over which the quantity is distributed and a function $\rho(\xi^a)$ such that:
 
-Define information entropy.
+\begin{equation}
+\int M \rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n = M
+\end{equation}
 
-Show change under coordinate transformation.
+We say that $\rho$ is the (normalized) density for our distribution and if we want to calculate the amount in a particular region $V$ we have:
 
-Show that conservation of information entropy is equivalent to transformation having Jacobian equals to 1.
+\begin{equation}
+m_V = M \mu(V) =\int_V M \rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n
+\end{equation}
 
-Show that canonical transformations are the ones and only the ones that have the Jacobian equals to one.
-Show relation to Poisson bracket.
-Show relation to incompressible flow.
-Show relation to symplectic form.
+The above expression would represent the mass or charge within a particular region, the amount of material of some selected set of compounds or the population in a particular age group.
+The integral is used when some of the variables $\xi^a$ are continuous and a simple sum would be used over the discrete ones.
 
-\section{Multiple degree of freedom}
+While the mathematical setting is the same for degrees of belief (credence) or the likelyhood that something will happen (probability), we will assume we are working with an actual physical distribution.
+Even when we are talking about marginal distributions or information entropy, we are doing it in the context of a system that has parts and each part has a different characterization.
 
-Show conceptually what needs to happen.
-Need to be able to define marginal distributions on each degree of freedom.
-Jacobian preserved under all degrees of freedom.
-Symplectic form in multiple dimension preserves area and orthogonality.
-Product of marginals needs to be preserved.
+The link to information theory is the following: suppose we have a distribution, known to both of us, and you select one element at random.
+How many yes or no questions do you have to answer before I know what case you selected?
+For example, if you have a pouch with 230 balls of which 115 are red and 115 are green, one yes or no question would be sufficient.
+In general, if $I$ is the number of questions then $C = 2 ^ I$ are the number of cases you can distinguish with those questions.
+But suppose $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
+That is, the normalized distribution is be the inverse of the number of cases: $\rho(\xi_0) = \frac{1}{C}$.
+Therefore the information needed to identify a particular element is $I(\xi^a)=\log \frac{1}{\rho(\xi^a)}$.
+If we take the average and change the base of the logarithm, we have:
+\begin{equation}
+I(\rho(\xi^a)) = \int \rho(\xi^a)) \ln \left(\frac{1}{\rho(\xi^a))}\right) d\xi^1 d\xi^2 ... d\xi^n =-\int \rho(\xi^a)) \ln (\rho(\xi^a)) d\xi^1 d\xi^2 ... d\xi^n
+\end{equation}
+which is the information entropy of the distribution.\footnote{The nature of information entropy is often misunderstood, especially in physics.
+It does not represent the information the distribution has: it represent the \emph{additional} information that would be needed to identify an element.
+Also, it does not represent knowledge: as we are studying actual physical distributions, any correct distribution has full knowledge of the system.}
 
-Show that setting up the problem mathematically is more problematic because: symplectic geometry does not have the tools to talk about distributions as it focuses on objects that are coordinate independent, statistics does not care that distribution and information entropy is coordinate dependent since the prime object is the cumulative distribution function.
+Technically, given the more convenient choice of the natural logarithm, the unit of information is the nat, which is equal to $\frac{1}{\ln 2}$ bits.
+However, in the discussion we will refer to it as measured in bits, as we believe that the majority of the readers is more conformable with the latter unit.
+Note that for continuous variables the expression has a slightly different meaning: the uniform distribution in a unit volume is assigned zero entropy so $I(\rho(\xi^a))$ represents the relative number of bits to achieve the same level of precision, which can be negative.
 
-Spend the rest at making links between geometric objects and distributions.
+\subsection*{Densities and change of variables}
 
-Show the definitions for mutual information entropy.
+There is one feature about densities we need to be fully aware of: they are not invariant under change of variables.
+Intuitively, if we change units of the denominator the distribution changes value.
+For example, $1$ kg/m becomes $1000$ kg/km: the number changes and the units allow us to understand that the overall quantity remains the same but mathematics does not keep track of units.
+In general, if $\hat{\xi}^a=\hat{\xi}^a(\xi^b)$ we can recover how the density changes by requiring that the total in each volume is the same no matter what variables are used.
+That is:
+\begin{align*}
+\int_V \rho(\xi^b) d\xi^1 ... d\xi^n &= \int_V \rho(\hat{\xi}^a) d\hat{\xi}^1 ... d\hat{\xi}^n \\
+&=\int_V\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\xi^1 ... d\xi^n
+\end{align*}
+\begin{equation}\label{density_transformation}
+\rho(\xi^b) = \rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|
+\end{equation}
+the density is multiplied by the absolute value of the Jacobian determinant of the transformation.
+As the Jacobian is in general a function of the variables themselves, this also means that whether the density at one point is greater or smaller than at another also depends on the choice of variables.
+
+Similarly, we can see that information entropy is not invariant
+\begin{align*}
+I(\rho(\xi^b)) &=-\int \rho(\xi^b) \ln (\rho(\xi^b)) d\xi^1 ... d\xi^n \\
+&=-\int \rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| \ln \left(\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|\right) d\xi^1 ... d\xi^n \\
+&=-\int \rho(\hat{\xi}^a) \ln \left(\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|\right) d\hat{\xi}^1 ... d\hat{\xi}^n \\
+&=-\int \rho(\hat{\xi}^a) \ln (\rho(\hat{\xi}^a)) d\hat{\xi}^1 ... d\hat{\xi}^n -\int \rho(\hat{\xi}^a) \ln \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\hat{\xi}^1 ... d\hat{\xi}^n
+\end{align*}
+\begin{equation}\label{entropy_transformation}
+I(\rho(\xi^b)) =I(\rho(\hat{\xi}^a)) -\int \rho(\hat{\xi}^a) \ln \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\hat{\xi}^1 ... d\hat{\xi}^n
+\end{equation}
+The reason is that we may be changing scale and therefore level of precision.
+
+The situation is actually worse.
+We'd like to think of a density as a number associated to a point of a space but that's technically incorrect.
+Given a manifold $\mathcal{M}$ upon which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \to \mathbb{R}$.
+That would be invariant change of variables, as a point is a point no matter how is it identified.
+Instead we additionally need a set of functions $\xi^b : \mathcal{M} \to \mathbb{R}$ and only then we can define the density as $\rho : \mathbb{R}^n \to \mathbb{R}$.
+The density value is defined on the values for a set of variables, not at a point.\footnote{As the density is really a limit, we need to know how that limit is taken and the coordinates decide that.} That is, the density is undefined without a (differentiable) set of coordinates.
+
+\subsection*{Mathematical tools for distributions}
+
+Different fields in math deal with this issue in different ways depending on their aims, which unfortunately leads to a fractured physical understanding given that one needs ideas and results from the different areas.
+In statistics and information theory one simply accepts the transformation rules and does not try to create invariant objects.
+This gives us the proper setting to study the statistical properties but it does not help us understand which mathematical elements represent objective physical entities.
+
+In probability theory one starts with three objects: a set of outcomes $\mathcal{M}$, a set of events $\sigma_\mathcal{M}$ where each event is a collection of outcomes, and a probability measure $\mu : \sigma_\mathcal{M} \rightarrow [0,1]$ that assigns a number between zero and one to each event.
+That is, we don't assign a probability density to the points of $\mathcal{M}$ but we assign a finite probability to sets of points.
+We can then define a random variable $\xi : \mathcal{M} \rightarrow \mathbb{R}$ as a real valued function of the outcomes.
+For each random variable we can define its cumulative distribution function $F_\xi(\xi_0)=\mu(\xi<\xi_0)$ and the probability density function is its Radon-–Nikodym derivative $\rho(\xi) = \frac{dF_\xi}{d\xi}$.
+That is: the integral is the primal object while the density is derived.
+
+This approach actually makes more physical sense: what we measure is the amount of mass in a finite region and the density is the limit for a smaller and smaller ones.
+The fundamental objects (outcomes, events and probability) are all defined without reference to variables.
+But in probability theory we have no notion of coordinate systems, vectors and all other geometrical notions.
+
+In differential geometry, one starts with a manifold: a set of points $\mathcal{M}$ which can be given local coordinates $\xi^i : \mathcal{M} \to \mathbb{R}$.
+One defines a tangent space $\mathsf{T}\mathcal{M}$ where vectors live and cotangent space $\mathsf{T}^*\mathcal{M}$ were linear functions of vectors live.
+Then we define $n$-forms as multi-linear functions of $n$ vectors that return the value associated with the infinitesimal parallelepiped they form.
+That is $\nu : (\mathsf{T}^*\mathcal{M})^n \rightarrow \mathbb{R}$.
+As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $m_V=\int_V \nu(dV)$ with no reference to coordinates.
+If $e_i$ are the basis vector associated with $\xi^i$ and $e^j$ are linear functions such that $e^j(e_i)=\delta_i^j$, then we can express each $n$-form as $\rho(\xi^i)e_1\wedge e_2 \wedge ...
+\wedge e_n$.
+The density $\rho(\xi^i)$ is the component of the $n$-form expressed in the $\xi^i$ coordinates in the a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value.
+But negative determinants means changing orientation of the region of integration, so fixing that would bring another sign change and $\rho(q^i)$ would not change.
+This is equivalent to saying that some coordinate systems are right handed and some are left handed.}
+
+This approach gives us a way to understand which objects are coordinate independent and which aren't.
+But it has two issues: it does not allow to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.\footnote{Typically, a vector is defined as a map between a scalar function of the manifold to another scalar function to the manifold.
+Or as an equivalence class of trajectories.
+It is a stretch to think of fluid velocities or electric fields in those terms.}
+
+So the situation is that we have a single clear physical objects we want to study, but different and somewhat disconnected mathematical frameworks to study it.
+This should not be surprising as the frameworks are defined by mathematicians to solve their own needs, which is totally understandable.
+But it is an issue as each mathematical framework only partially captures our physical understanding and we risk missing the full picture.
+
+\section{Coordinate independent densities}
+
+Suppose we have a system which we can think of made of infinitesimally small parts, which we call particles.
+Suppose  that each infinitesimal part can be assigned a state from a state space $\mathcal{S}$.
+The state of the whole system will be a distribution over those states.
+Now suppose that this system evolves in a way that is deterministic and reversible, that is the state of a particle at one time is associated with one and only one state at another time.
+What does that tell us about our distribution?
+
+If the evolution is deterministic and reversible, then all the parts that are in one state will be mapped to another: the density should be exactly mapped to a different state and therefore should not change value.
+If the evolution is deterministic and reversible, then if I give you enough information to identify an element at one time I also gave you enough information to identify the element at a future time: the information entropy will remain the same over time.
+And here lies the problem: how can we make sure that these quantities are preserved over time if they are not even observer independent?
+How can we know if the density stays the same if we can't even compare the value of the densities at two different elements of the distribution?
+How can we tell whether the information entropy is conserved if we can't event define a unique value at a given time?
+
+\textbf{If we want to define deterministc and reversible evolution over distributions, they must be invariant distributions.
+That is, the densities and information entropy must be independent on the choice of coordinates.}
+
+If our variables are discrete, all our integrals are sums and all our definitions work.
+Everything is already coordinate invariant.
+If our variables are continuous, instead, we have a problem.
+Because the densities and the information entropy vary with the Jacobian, we need something more.
+Our state space $\mathcal{S}$ must allow to write densities whose value does not change under coordinate transformation.
+This seems to be impossible given what we saw before: how can it work?
+
+Suppose $\mathcal{S}$ is a two dimensional manifold.
+Suppose we have two variables $(q,k)$ such that $k$ uses the inverse units of $q$.
+For example, if $q$ is meters then $k$ is inverse meters and an infinitesimal area $dq dk$ will be a pure number.
+We say $q$ is a coordinate as it is a variable that defines a unit, while $k$ is the corresponding conjugate variable as it uses the inverse unit.
+Now suppose you change coordinate.
+For example, $\hat{q}$ will be kilometers and $\hat{k}$ inverse kilometers.
+We'll have $d\hat{q} d\hat{k} = dq dk$.
+Therefore we'll also have $\rho(q,k) = \rho(\hat{q}, \hat{k})$: the density is invariant and so is the information entropy.
+
+If we have a set of variables fixed by a single choice of unit then what we discussed before is the only possible way.
+Let $q$ be the coordinate, the variable that defines the unit.
+Suppose we have $n$ variable $\xi^a$ such that $\xi^1 = q$.
+Each vector $v^a$ tangent to that space has $n$-components.
+If we change unit $\hat{q}=\hat{q}(q)$ we have two constraints: the one unit change and the unitary Jacobian, and these must be enough to determine how the vector components change.
+The components must be equal to the constraints and therefore they are exactly two variable $\xi^a = (q, k)$.
+Then $k$ is the conjungate variable of $q$.
+
+\subsection*{Hamiltonian mechanics for one degree of freedom}
+
+Let's set $\hbar$ as the unit we use to describe the amount of possible states. We have:
+\begin{equation}
+m_V = M \mu(V) =\int_V M \rho(q, k) \hbar dq dk = \int_V M \rho(q, p) dq dp
+\end{equation}
+where we set $p=\hbar k$. We should recognize $p$ as conjugate momentum and $k$ the classical analogue of the wave number. Suppose we change variables. In general the Jacobian will be given by:
+\begin{equation}
+\label{Poisson}
+\begin{aligned}
+|J| &= \left| \begin{matrix}
+\dfrac{\partial \hat{q}}{\partial q} & \dfrac{\partial \hat{q}}{\partial p} \\[2.2ex]
+\dfrac{\partial \hat{p}}{\partial q} & \dfrac{\partial \hat{p}}{\partial p} \end{matrix} \right| = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial x} \frac{\partial \hat{q}}{\partial p} &= \{\hat{q}, \hat{p}\}
+\end{aligned}
+\end{equation}
+which we recognize it to be the Poisson bracket. If the change of variable is simply a change of coordinates, a change of units, we have:
+\begin{equation}
+\label{coordinate_change}
+\begin{aligned}
+\hat{q} &= \hat{q}(q) \\
+\{\hat{q}, \hat{p}\} &= 1 = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} \\
+\dfrac{\partial \hat{p}}{\partial p} &= \frac{\partial \hat{q}}{\partial q} ^{-1} = \frac{\partial q}{\partial \hat{q}} \\
+\hat{p} &= \frac{\partial q}{\partial \hat{q}} p
+\end{aligned}
+\end{equation}
+Conjugate momentum transforms as a covariant component. Under coordinate transformations we have:
+\begin{equation}
+\label{density_invariance}
+\begin{aligned}
+\rho(q,p) &= \rho(\hat{q}, \hat{p}) \\
+I(\rho(q,p)) &= I(\rho(\hat{q},\hat{p}))
+\end{aligned}
+\end{equation}
+Let us call $(q,p)$ canonical variables if the density is equal to the invariant density.
+Then two other variables such that $\{\hat{q}, \hat{p}\}=1$ are also canonical since the Jacobian determinant will be unitary.
+
+We can think of the evolution as a vector field on $\mathcal{S}$ of components $S = (S_q, S_q) = (\frac{dq}{dt}, \frac{dp}{dt})$ that gives the direction in which states move in time.
+The Jacobian for time evolution will be
+\begin{equation}
+\label{Jacobian_evolution}
+\begin{aligned}
+|J| &= \left| \begin{matrix}
+1 + \dfrac{\partial S_q}{\partial q}dt & \dfrac{\partial S_q}{\partial p} dt \\[2.2ex]
+\dfrac{\partial S_p}{\partial q}  dt & 1 + \dfrac{\partial S_p}{\partial p} dt \end{matrix} \right| \\
+&= 1 + \left[ \dfrac{\partial S_q}{\partial q} + \dfrac{\partial S_p}{\partial p} \right]dt + O(dt^2)\\
+&= 1 + div(S)dt + O(dt^2)\\
+\end{aligned}
+\end{equation}
+
+If we want the time evolution to transport the densities and to conserve information entropy, the Jacobian must be unitary and therefore $S$ is divergence-free and admits a potential. We have
+\begin{equation}
+\label{Potential_Hamilton}
+\begin{aligned}
+S = - curl(H) &= (\frac{\partial H}{\partial q}, - \frac{\partial H}{\partial p}) \\
+\frac{dq}{dt} &= \frac{\partial H}{\partial p}  \\
+\frac{dp}{dt} &= - \frac{\partial H}{\partial q}  \\
+\end{aligned}
+\end{equation}
+and we recognize Hamilton's equations.
+Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it's not just that the density and the areas are conserved in time: is that they are the same for all canonical coordinates.
+That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is the exactly the evolution of (classical) distributions that preserve information entropy and densities}.
+
+We'd like to stress how all elements of Hamiltonian mechanics had a straight forward physical meaning in terms of invariant densities and change of coordinates and variables.
+
+\subsection*{Distributions over multiple degrees of freedom}
+
+To understand what happens in multiple degrees of freedom, let's first look at change of variables of marginal distributions.
+Suppose we have a distribution $\rho(\xi^a)$ that depends on many coordinates, on many units.
+If the density is coordinate invariant, then it will be so even if we change one coordinate, say $\xi^1$.
+As we saw before, then we can find the conjugate variable $\xi^2$ that changes in the opposite way.
+We can calculate the marginal distribution by integrating over those two quantities.
+\begin{equation}
+\rho(\xi^b) = \int \rho(\xi^a) d\xi^1 d\xi^2
+\end{equation}
+where $\xi^b$ are the remaining coordinates.
+If the variables $\xi^b$ are independent from $\xi^1$ and $\xi^2$, then we must be able to change units independently as well.
+Which means $\rho(\xi^b)$ must be coordinate invariant.
+
+We can imagine to proceed iteratively, one unit at a time.
+We'll find that our variables are paired $(q^i, p_i)$  with the first defining the units and the seconds defining the invariant areas.
+We call each pair an independent degree of freedom as it constitute an independent choice of unit.
+We recognize $\mathcal{Q}$, the manifold charted by $q^i$, as the (poorly named) configuration space.
+We are going to call it coordinate space as it includes all variables and only the variables that define the units and the reference frame.
+The state space for the particles, then, is a set of coordinates plus an equal number of conjugate variables that vary like covector components, therefore we recognize $\mathcal{S}=\mathsf{T}^*\mathcal{Q}$ as the cotangent bundle.\footnote{This does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number.
+The state space is just a manifold and the structure it has comes from how units transform, not because the variable represent different objects.
+Unfortunately mathematical structures strip out the units, so that knowledge is lost and then resurfaces in other structures.
+Changing state variables from position and momentum to position and velocity is not, as one may be mistakenly understand from the math, a change of objects: is relabeling the same states much like changing spatial coordinates relabels positions.
+It is not, in general, a canonical transformation as the density will not correspond to the invariant one, and it is not simply a change of coordinates.
+But it is still a valid change of state variables.}
+
+In terms of information theory, this is equivalent to being able to assign a coordinate invariant information entropy to each of the marginal distributions.
+This means that relationships between the entropy of the total and marginal distributions are preserved.
+For example, saying that the marginal distributions $\rho_2(q^2)$ and $\rho_1(q^1)$ are independent is equivalent to saying that the entropy of the joint distribution is equal to the sum of the entropy of the marginals: $I(\rho(q^1,q^2)) = I(\rho_2(q^2)) +I(\rho_1(q^1))$.\footnote{This is also equivalent to saying that the joint distribution is the product of the marginals $\rho(q^1,q^2) = \rho_2(q^2)\rho_1(q^1)$.} That is, giving both values is equivalent to giving one and then the other.
+This means we can think of these relationships and quantities as real physical properties assigned to the degree of freedom and not just to the coordinates.
+
+\subsection*{The geometry of phase space}
+
+Now that we have an intuitive sense of the objects we want to describe physically, we can turn to differential geometry to derive the geometrical structure of the state space and then the equations of motion.
+What we want to capture is that we can integrate densities over two dimensional sub-manifolds in a way that is coordinate invariant.
+That is, we need to have a two-form $\omega$ such that:
+\begin{equation}
+\rho_\Sigma = \int_\Sigma \rho(\xi^a) \omega(d\Sigma)
+\end{equation}
+is coordinate invariant.
+In fact, creating a marginal distribution means giving a family of surfaces parameterized by a set of variables, that do not intersect and whose union is the whole space and then calculating the integral as a function of the variables.
+
+The requirement that $\omega$ is a two-form, i.e.
+linear and anti-symmetric, comes from the fact that it needs to represent areas.
+That is, it needs to be linear because the area of a parallelogram is a linear function of its sides.
+And if $v^1$ and $v^2$ are vectors and $\omega(v^1, v^2)$ represents the area, then rotating by 90 degrees will give the same area.
+That is, $\omega(v^1, v^2) = \omega(v^2, -v^1) = -\omega(v^2, v^1)$.
+Therefore it must be anti-symmetric.
+It should also be non-degenerate, i.e.
+there is no direction for which $\omega$ is always zero or we would not be able to define a volume.
+
+We also want the two-form to be closed, that is the integral
+\begin{equation}
+\int_\Sigma \omega(d\Sigma)
+\end{equation}
+is zero if $\Sigma$ is a closed surface.
+To understand why, note that it corresponds to the case where the density is the unit constant.
+If the surface is a parallelopiped in some variables, each face that give a positive contribution will have a parallel and equal face that would give an opposite contribution.
+As the closed integral needs to be zero for all parallelopipeds, it will also be zero for an arbitrary closed surface.
+
+Under a change of variables, the two-form $\omega$ must be invariant.
+That is, phase space is a symplectic manifold.
+And, as any symplectic manifald, we can find Darboux variables\footnote{In general, this gives a general understanding of what a differentiable structure is.
+Any manifold, even a non-differentiable one, can identify points with coordinates.
+A differentiable structure on a manifold selects a set of charts that allow differentiable coordinate change.
+But there are many such sets, which one do we pick?
+We pick the one that allows us to write densities for our distributions.
+Similarly, a symplectic manifolds selects only the coordinates that allows us to write invariant densities.
+Also note how both the tangent space and the space of distributions are vector spaces.
+In a way, the linearity of distributions corresponds to the linearity of the tangent space.} $q^i, p_j$ and express the two-form as
+\begin{equation}
+\label{Symplectic}
+\begin{aligned}
+\omega &= dq^i \wedge dp_i = d\xi^a\omega_{ab}d\xi^b \\
+\omega_{ab} &=  \left[
+\begin{array}{cc}
+0 & 1 \\
+-1 & 0 \\
+\end{array}
+\right] \otimes I_n =
+\left[
+\begin{array}{cc}
+0 & I_n \\
+-I_n & 0 \\
+\end{array}
+\right]
+\end{aligned}
+\end{equation}
+
+The components $\omega_{ab}$ let us understand what geometrical features are captured by $\omega$.
+It corresponds to a two dimensional vector product within each degree of freedom while it corresponds to a scalar product across degrees of freedom.
+As $\omega$ is preserved during coordinate transformations, the area within each degree of freedom is preserved meaning we are preserving the marginal distribution as well.
+Across degree of freedom, instead, we are preserving the angle meaning two perpendicular degrees of freedom remain perpendicular.
+But here perpedicular means independent: the density within the four dimensional volume is the product of the densities on sides.
+As these are preserved under coordinate transformation, these now can correspond to actual physical entities.
+
+\subsection*{Hamiltonian mechanics for multiple degrees of freedom}
+
+Deterministic and reversible evolution will preserve the densities and therefore preserve $\omega$.
+That is, $\omega(v, w) = \omega(v',w')$ where $v'$ and $w'$ are the vector evolved in time.
+In fancy math terms, deterministic and reversible evolution for a distribution is a symplectomorphism.
+As before, $S = \left(\frac{d\xi^a}{dt}\right)$ is the vector field corresponding to the evolution.
+The vector components change according to $v'^a = \partial_b \xi(t+dt)^a v^b = (\delta^a_b + \partial_b S^a dt) v^b$.
+Therefore we have:
+\begin{align*}
+v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
+&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
+&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} + v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt \\ &+ O(dt^2)
+\end{align*}
+If we set $S_{b} \equiv S^{a} \omega_{ab}$, we have:
+\begin{equation}
+\begin{aligned}
+v^{c} w^{b} \partial_{c} S_{b} &- v^{a} w^{d} \partial_{d} S_{a} = 0\\
+\partial_{a} S_{b} - \partial_{b} S_{a} &= curl(S_{a}) = 0 \\
+S_{a} &= \partial_{a}H \\
+\frac{dq^i}{dt} &= \frac{\partial H}{\partial p_i}  \\
+\frac{dp_i}{dt} &= - \frac{\partial H}{\partial q^i}
+\end{aligned}
+\end{equation}
+
+We recognize Hamilton's equations for multiple degrees of freedom.
 
@@
 
-Why should we care? Because it provides links.
+Hamiltonian mechanics, through Liouville's theorem, has always had an important role in statistical mechanics.
+What we are showing is that that role is not a coincidence: it is the very defining characteristic.
+\textbf{Hamiltonian mechanics is a type of statistical mechanics}, one with no probability and deterministic and reversible laws.
+The fact that entropy can be defined objectively and is conserved through evolution cannot be simply seen as a curiosity.
 
-Uncertainty principle as relationship between entropy and uncertainty.
+To us, it does not make sense to start our physical discourse by imposing a state space for point-like particles, that somehow does not require only points but a particular differentiable structure as well.
+Then, if we are so keen, put distributions on it which turn out, somewhat magically, to have density and entropy defined in a coordinate independent way.
+It makes a lot more sense to start by saying we have a finite system which we pretend to be made of infinitesimal parts, which we keep to dividing as we please and derive the state space of particles as the limit of such process of subdivision.
+This gives uniquely gives us the actual mathematical structure and its physical meaning.
 
-Relativistic Hamiltonian motion as preservation of entropy.
+Not only the second approach is preferable because with fewer starting points it justifies the state space, the evolution and why that state space is so compatible with statistical mechanics.
+It also maps better to what we do in classical physics: we study composite systems (planets, cannonballs, beads on a wire, fluids) which we sometime pretend we can recursively divide.
+None of those objects, in fact, has a single position or momentum: only the center of mass has a single position and momentum but those are precisely the expectation values $\int q \rho(q,p) dq dp$ and $\int p \rho(q,p) dq dp$ over the distribution.
+In other words, particles are not point-like in classical mechanics.
+They are infinitesimal cells of phase-space and that's why they evolution is symplectomorphism: that's why phase space requires a differentiable structure.
 
-Relationship to conservation of energy in isolated systems.
+Under this light, asking whether classical mechanics is tenable coincides with asking whether we can really keep dividing our object indefinitely.
+If one is convinced that that is not realistic then he is already convinced that classical mechanics will not hold in those regimes.
+And if that simplification fails, it's not that we need an altogether new set of rules: the macroscopic object is still the same.
+We simply need to change how the subdivision works.
+
+While the power of this shift can really be appreciated in the context of our broder work, it should be clear how with relative few starting points, elements of relativity (e.g.
+coordinate invariance), statistical mechanics (e.g.
+physical distributions), quantum mechanics (e.g.
+wave number, conjugate quantities) and classical Hamiltonian mechanics (e.g.
+phase space, Poisson brackets) already emerge and form a coherent picture.
+In particular note how energy (in the form of the Hamiltonian) and entropy (in the form of information entropy) both come out from the single unassuming idea of invariant distributions.
+And note how conservation of entropy is related to deterministic and reversible evolution.
+Also note how deterministic and reversible evolution means the that the system is only influenced by itself, which means it is not influenced by anything else, which means it is isolated, which indeed means energy is conserved.
+This surely this cannot be all coincidence.
+
+The larger point is that physics is one and it can't be understood piecemeal.
+Nature does not separate between dynamical systems and thermodynamics, between topologies and $\sigma$-algebras, between classical systems and quantum systems.
+The boundaries between all these disciplines are as much unfortunate as they are artificial.
+There is no real understanding of one without the other.
+
+Having a coherent picture across the different areas not only is intellectually more satisfying, but it allows us to switch perspective when the intuition from one fails.
+As a practical example, we can show how understanding the link between Hamiltonian mechanics and information theory leads very straightforwardly to a classical equivalent for the uncertainty principle.
+
+\subsection*{Classical uncertainty principle}
+
+As we saw, Hamiltonian dynamics not only conserves the energy of all particles, but it also conserves the information entropy of the distribution as a whole.
+We can imagine that there is going to be some link between the spread of a distribution and the number of bits required to identify an element.
+So we ask: what is the distribution that minimizes the spread given a certain amount of information entropy?
+If $I_0$ is the set amount of information entropy and $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$ is the spread, we can use Lagrange multipliers to find out.
+\begin{align*}
+L = &\int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp \\
+&+ \lambda_1(\int \rho dqdp - 1) \\ &+ \lambda_2(- \int \rho \ln \rho \, dqdp - I_0)\\ 
+\delta L = &\int \delta \rho [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho - \lambda_2 ] dqdp = 0 \\
+\lambda_2 \ln \rho = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
+\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}\\
+\end{align*}
+We solve the multipliers and have:
+\begin{align*}
+\rho = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
+I_0 = &\ln (2\pi\sigma_q\sigma_p) + 1
+\end{align*}
+The distribution that minimizes the spread, then, is the product of two independent Gaussians.
+As the entropy is conserved during Hamiltonian evolution the product $\sigma_q^2 \sigma_p^2$ can never be less than the one given by the Gaussian distribution of the same entropy.
+We have:
+\begin{align*}
+\sigma_q\sigma_p \geq \exp (I_0 - 1) / 2 \pi 
+\end{align*}
+Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and their elements, and therefore it can't shrink it more than a set amount.
+If it did, it would start concentrating the density, meaning that multiple states have part of their density mapped to a single state and the dynamics would not be reversible.
+
+\section{Conclusion}
+
+In this paper we have shown how classical Hamiltonian mechanics coincides with conservation of information entropy.
+The structure of phase space is required to be able to define densities and entropy in a way that is coordinate invariant, and therefore physically meaningful.
+This clarifies the link between Hamiltonian mechanics, statistical mechanics and deterministic and reversible motion.
+
+In the narrower context, it allows to understand what physical object is represented by each mathematical one. This can be summarized in table \ref{dictionary}.
+\begin{table}[h]
+	\centering
+	%	\begin{tabular}{p{0.2\textwidth} p{0.1\textwidth} p{0.1\textwidth} p{0.5\textwidth}}
+	\begin{tabular}{c p{0.3\textwidth} p{0.5\textwidth} }
+		& Name & Meaning\\ 
+		\hline 
+		& Particle (classical) & an infinitesimal part of the system (i.e. the limit of recursive division) \\ 
+		$\mathcal{S} =\mathsf{T}^*\mathcal{Q}$ & Phase space \newline (Tangent bundle) & the set of all possible states for particles upon which invariant distributions can be defined \\
+		$\xi^a$ & State variables \newline (Unified coordinates) & the set of variables needed to identify the state of a particle \\ 
+		$\rho(\xi^a)$ & Density & the amount of material for a given particle state (i.e. in the limit of the recursive division)\\ 
+		$I(\rho(\xi^a))$ & Information entropy & the number of bits to identify an element of the distribution to a unitary level of precision\\ 
+		$q^i$ & Coordinate & a state variable that also defines a unit \\
+		$\mathcal{Q}$ & Coordinate space \newline (Configuration space) & the space charted by all coordinates \\
+		$k_i$ & Conjugate coordinate & a state variable that uses the inverse unit of the corresponding coordinate \\
+		$\hbar$ & & the unit for measuring the range of possible states within a single degree of freedom \\
+		$p_i=\hbar k_i$ & Conjugate momentum & a state variable that together with the corresponding coordinate defines invariant ranges of possible states \\
+		$(q^i, p_i)$ & Canonical variables \newline (Canonical coordinates) & a set of variable for which the density is invariant and expressed over units of $\hbar$\\ 
+		& Canonical transformation & a change in state variables that does not change the density\\ 
+	\end{tabular}
+	\caption{Dictionary between mathematical and physical objects.}
+	\label{dictionary}
+\end{table}
+It also allows us to see that deterministic and reversible evolution must conserve entropy and that isolated systems must conserve energy.
+In the broader context, it allows to make deeper connections between different areas of math and physics, and pushes us to rethink our starting points.
+In particular, it tells us that classical particles should not be considered point-like objects (upon which entropy and densities cannot be defined), but the limit of recursive subdivisions (i.e.
+infinitesimal cells of phase space).
+
+These insights come from our project, Assumptions of Physics, that aims to re-derive the standard laws from a handful of physical starting points.
+We believe that a reorganization of fundamental physics on more physically meaningful conceptual footing and more rigorous mathematical grounds is not only possible, but long overdue.
+The insights presented here are one step toward that direction as we know them to be still valid and useful in a larger setting, such as understanding the similarities in the mathematical structures of classical and quantum mechanics.
 

### 2018-10-26 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
First draft probably done

@@
 \begin{abstract}
-	In this work we show that the canonical transformations of classical Hamiltonian mechanics are exactly the transformations that preserve information.
+	In this work we show the equivalence between Hamiltonian mechanics and conservation of information entropy.
+We will show that distribution with coordinate independent values for information entropy require that the manifold on which the distribution is defined is charted by conjugate pairs (i.e.
+it is a symplectic manifold).
+We will also show that further requiring that the information entropy is conserved during the evolution yields Hamilton's equations.
+This will provide new insights in the nature of classical mechanics and physics in general, as it will connect to elements of relativity, statistical mechanics, thermodynamics, quantum mechanics, probability theory and symplectic geometry.
+This work is part of a larger project, called Assumptions of physics, which aims to rederive the known equations from a handful of physically meaningful starting points.
 \end{abstract}
@@
 \section{Introduction}
+
+Over the past decade there has been a renewed interest in the underpinning of classical mechanics~\cite{North,Curiel,Barrett1,Barrett2}.
+Different attempts and arguments have been put forward to understand what are the most important feature of the different mathematical formulations, whether one is more fundamental of the other in some sense or whether they are equivalent.
+
+All these attempts, though, suffer from one problem.
+They concentrate on the mathematical structure.
+The issue is that, in general, a mathematical object can represent more than one physical entity.
+For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.
+On the other end, the same physical entity can be described by more than one mathematical object.
+For example, the same cannonball can be described by position and velocity in a Lagrangian setting, position and momentum in a Hamiltonian setting or in a thermodynamics setting by temperature, volume and so on.
+The connection between math and physics, then, is many-to-many and therefore we can't study only one side.
+The reason is that mathematics captures only the syntax of a particular description without its semantics, which at that point cannot be recovered.
+Mathematical structures are just shadows of the physical model they represent.
+
+It should be clear that we do not fault the authors.
+Even just a century ago, there was a push to find fundamental principles or laws that could serve as a starting point for the different theories.
+Newtonian mechanics, thermodynamics and special relativity serve as good examples.
+Nowadays, unfortunately, much of the current fundamental theories, including Lagrangian, Hamiltonian, Quantum mechanics and Quantum Field Theory take as their starting point a particular mathematical structure.
+That is, we do not know what the value of a Lagrangian is or what the symplectic form represents: we just take them as a given.
+Moreover, a lot of work in theoretical physics nowadays starts by exploring novel abstract mathematical structures and then trying to see whether some physical meaning or prediction can be extracted.
+We are not going to debate here how and why this shift happened.
+We just merely note that it leaves the semantics of our theories ill defined.\footnote{Which is a nice way to say that we don't really know what are we talking about.}
+
+It is only natural that, given the circumstances, one uses the only thing that seems to be well defined, the math, and tries to get understanding from there.
+But here lies the fundamental problem: that math is not created by physicists to do physics, it is created by mathematicians to do math.
+The definitions of topology, differentiable manifold, symplectic manifold, Hilbert space and so on are chosen either because they are convenient to prove theorems or because they are convenient to make calculations.\footnote{In fact, it is common in math to have two such definitions and then prove they are equivalent.} We should not expect them to be a good match for definitions that would be physically meaningful.
+In other words, there is a very good chance that just looking at just the math may lead us, like horses with blinders, to a very clear path which is not the physically meaningful one.\footnote{To be clear, we do not fault the mathematicians either as they are not responsible for whether their structures are used appropriately or not.}
+
+Fully conscious of these issues, we have embarked in an effort, called Assumptions of Physics~\cite{carc1}, that aims to find a handful of physical principles from which the known laws of physics can be formally derived~\cite{carc2}.
+The approach has several advantages.
+First, by starting from physical ideas we put physics back at the center of the discussion, and not relegate it as a mere after the fact interpretation: the semantics is always well defined.
+Second, it puts physics on sturdier mathematical grounds as we have a precise physical meaning to all the formal objects and we leave out those that do not.
+Third, it fosters connections between different fields of science, math, philosophy and engineering: it is our belief that nature is one and does not care about such divisions and therefore truly fundamental concepts must span freely across different areas.
+Fourth, it gives a framework to pose deep questions of the type the philosophy of science community is interested in.
+And lastly, by being forced to formalize and make explicit our assumptions, we can better understand how and why they fail possibly giving new insights that may lead to new physics.
+While the project is ambitious, we have already made progress beyond what we thought it could be reasonably done.
+So there's that.
+
+In this paper we focus on just one particular insight provided by our work: \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
+a symplectic structure where variables are organized in conjugate pairs).
+In section 2 we show that, in general, information entropy and distributions densities are coordinate dependent.
+In section 3 we show that the structure of phase space is required to make such objects invariant under coordinate transformation (i.e.
+observer independent) and therefore give them an objective physical character.
+Conservation of entropy over time will force that invariance during time evolution, giving rise to Hamilton's equations.
+In chapter 4 we will briefly discuss the result and show how it necessarily leads to a classical version of the uncertainty principle.
+
+The overall idea is that \textbf{Hamiltonian mechanics is best understood as the deterministic and reversible\footnote{By reversible we mean the ability, given a state, to reconstruct its past.
+This is the dual of determinism, that is the ability, given a state, to predict its future.} evolution of infinitesimally reducible objects}.
+As the information needed to identify an element of the distribution does not change in time, knowing the initial state is equivalent to knowing the future or past state.
+Classical particles, then, should be understood as the limit of subsequent subdivisions, infinitesimally small objects with extent in phase space.
+This single physical model is captured piecemeal by elements of statistics, probability theory and symplectic geometry which are given a coherent and well defined meaning.
+In our view, this understanding also meshes better with other ideas from general relativity\footnote{A point particle in general relativity would not follow a geodesic and would, due to the infinite mass density, puncture space-time.}, thermodynamics and quantum mechanics.
 
@@
 How many yes or no questions do you have to answer before I know what case you selected?
-For example, if you have a pouch with 230 balls of which 115 are red and 115 are green, one yes or no question would be sufficient.
+For example, suppose there are 4 balls of which 2 are red and 2 are green, one yes or no question would be sufficient.
 In general, if $I$ is the number of questions then $C = 2 ^ I$ are the number of cases you can distinguish with those questions.
-But suppose $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
-That is, the normalized distribution is be the inverse of the number of cases: $\rho(\xi_0) = \frac{1}{C}$.
+On the other hand, suppose $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
+That is, the normalized distribution is the inverse of the number of cases: $\rho(\xi_0) = \frac{1}{C}$.
 Therefore the information needed to identify a particular element is $I(\xi^a)=\log \frac{1}{\rho(\xi^a)}$.
@@
 The reason is that we may be changing scale and therefore level of precision.
+Given that both change is determined by the Jacobian, we conclude that \textbf{during a change of variable the information entropy of all possible distributions does not change if and only if the density remains the same at every point under the same change of variable}.
 
-The situation is actually worse.
-We'd like to think of a density as a number associated to a point of a space but that's technically incorrect.
+To sum up, while we'd like to think of a density as a number associated to a point of a space, that's technically incorrect.
 Given a manifold $\mathcal{M}$ upon which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \to \mathbb{R}$.
-That would be invariant change of variables, as a point is a point no matter how is it identified.
-Instead we additionally need a set of functions $\xi^b : \mathcal{M} \to \mathbb{R}$ and only then we can define the density as $\rho : \mathbb{R}^n \to \mathbb{R}$.
+That would be a scalar function, invariant under change of variables, as a point is a point no matter how is it identified.
+Instead we additionally need a set of functions $\xi^a : \mathcal{M} \to \mathbb{R}$ and only then we can define the density as $\rho : \mathbb{R}^n \to \mathbb{R}$.
 The density value is defined on the values for a set of variables, not at a point.\footnote{As the density is really a limit, we need to know how that limit is taken and the coordinates decide that.} That is, the density is undefined without a (differentiable) set of coordinates.
@@
 In probability theory one starts with three objects: a set of outcomes $\mathcal{M}$, a set of events $\sigma_\mathcal{M}$ where each event is a collection of outcomes, and a probability measure $\mu : \sigma_\mathcal{M} \rightarrow [0,1]$ that assigns a number between zero and one to each event.
-That is, we don't assign a probability density to the points of $\mathcal{M}$ but we assign a finite probability to sets of points.
+That is, we don't assign a probability density to the points of $\mathcal{M}$ but we assign a finite probability to finite regions.
 We can then define a random variable $\xi : \mathcal{M} \rightarrow \mathbb{R}$ as a real valued function of the outcomes.
@@
 That is $\nu : (\mathsf{T}^*\mathcal{M})^n \rightarrow \mathbb{R}$.
-As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $m_V=\int_V \nu(dV)$ with no reference to coordinates.
-If $e_i$ are the basis vector associated with $\xi^i$ and $e^j$ are linear functions such that $e^j(e_i)=\delta_i^j$, then we can express each $n$-form as $\rho(\xi^i)e_1\wedge e_2 \wedge ...
+As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $\mu(V)=\int_V \nu(dV)$ with no reference to coordinates.
+If $e_a$ are the basis vector associated with $\xi^a$ and $e^b$ are linear functions such that $e^b(e_a)=\delta_a^b$, then we can express each $n$-form as $\rho(\xi^a)e_1\wedge e_2 \wedge ...
 \wedge e_n$.
-The density $\rho(\xi^i)$ is the component of the $n$-form expressed in the $\xi^i$ coordinates in the a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value.
-But negative determinants means changing orientation of the region of integration, so fixing that would bring another sign change and $\rho(q^i)$ would not change.
+The density $\rho(\xi^a)$ is the component of the $n$-form expressed in the $\xi^a$ coordinates in the a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value.
+But negative determinants means changing orientation of the region of integration, so fixing that would bring another sign change and $\rho(\xi^a)$ would not change.
 This is equivalent to saying that some coordinate systems are right handed and some are left handed.}
@@
 Or as an equivalence class of trajectories.
-It is a stretch to think of fluid velocities or electric fields in those terms.}
+It is a stretch to think of fluid velocities or electric fields in those terms.
+For that same reason, we do not use the differential geometry notation $\frac{\partial}{\partial \xi^a}$ and $d\xi^a$ for vector and covector basis because they do not convey the actual physical meaning of these objects as they are used in physics.}
 
 So the situation is that we have a single clear physical objects we want to study, but different and somewhat disconnected mathematical frameworks to study it.
-This should not be surprising as the frameworks are defined by mathematicians to solve their own needs, which is totally understandable.
-But it is an issue as each mathematical framework only partially captures our physical understanding and we risk missing the full picture.
+This should not be surprising as the frameworks are defined by mathematicians, as we said before, to solve their own needs.
+But it is an issue when these are used in physics without careful consideration as each mathematical framework only partially captures our physical understanding.
+That is why we argue that simply examining the mathematical structure used in a physical theory is not enough to gain a complete picture.
 
-\section{Coordinate independent densities}
+\section{Coordinate independent distributions}
 
@@
 If the evolution is deterministic and reversible, then all the parts that are in one state will be mapped to another: the density should be exactly mapped to a different state and therefore should not change value.
-If the evolution is deterministic and reversible, then if I give you enough information to identify an element at one time I also gave you enough information to identify the element at a future time: the information entropy will remain the same over time.
+If the evolution is deterministic and reversible, then if I give you enough information to identify the state of an element at one time I also gave you enough information to identify the state of the same element at a future time: the information entropy will remain the same over time.
 And here lies the problem: how can we make sure that these quantities are preserved over time if they are not even observer independent?
@@
 
-\textbf{If we want to define deterministc and reversible evolution over distributions, they must be invariant distributions.
-That is, the densities and information entropy must be independent on the choice of coordinates.}
+This leads us to the following insight: \textbf{if we want to define deterministc and reversible evolution over distributions, they must be invariant distributions.
+That is, the densities and information entropy must be independent on the choice of coordinates.} As we noted before, requiring one to be coordinate invariant will automatically require the other to be invariant as well.
 
@@
 Suppose we have two variables $(q,k)$ such that $k$ uses the inverse units of $q$.
-For example, if $q$ is meters then $k$ is inverse meters and an infinitesimal area $dq dk$ will be a pure number.
+For example, if $q$ is meters then $k$ is inverse meters which means an infinitesimal area $dq dk$ will be a pure number.
 We say $q$ is a coordinate as it is a variable that defines a unit, while $k$ is the corresponding conjugate variable as it uses the inverse unit.
@@
 
+Physically, the coordinate plays a double role.
+It both defines a unit and it helps identify particle states.
+It is this double role that gives it an important character because a change in $q$ must induce a change in $k$ as their units and the unit of $\rho$ are not independent.
+Symplectic geometry strips away the units and does not elevate densities as primary objects.
+It is no wonder the link gets lost.
+
 \subsection*{Hamiltonian mechanics for one degree of freedom}
 
-Let's set $\hbar$ as the unit we use to describe the amount of possible states. We have:
+Let's set $\hbar$ as the unit we use to describe the amount of possible states (or the inverse of a unitary density). We have:
 \begin{equation}
@@
 
-We'd like to stress how all elements of Hamiltonian mechanics had a straight forward physical meaning in terms of invariant densities and change of coordinates and variables.
+We'd like to stress how, once the simple premise of invariant distributions is accepted, all elements of Hamiltonian mechanics are necessary and have a straight forward physical meaning.
+Not only that: because of the equivalent we know that everything in Hamiltonian mechanics can be understood with just that simple premise.
+And the premise is not about cotangent bundles of Lie group of quasi-q spaces on an infinite-dimensonal lattice.
+It's just the simple realization that if we require distributions to have a physically objective reality they must be coordinate independent.
 
@@
 
+The symplectic structure was recovered by requiring invariant distributions which in turns requires invariant integration over independent degrees of freedom.
+Note how the requirement was given outside of the symplectic structure itself and therefore it cannot be recovered simply from that.
+
 \section{Discussion}
@@
 
+Also note that this view tells us exactly what the mathematical framework of symplectic geometry is supposed to represent.
+The state space of particles is a manifold because we use real numbers to identify the states.
+Among all possible choice of state variables, we conveniently choose the ones in which the distribution can be expressed as a density.
+That's the differentiable structure on the manifold.
+The tangent and cotangent space are the tools to keep track of the process of infinitesimal subdivision and they are linear spaces because the space of all possible distributions is itself a linear space.
+Moreover the differentiable manifold has a symplectic structure because the density is invariant under coordinate transformation.
+Therefore among all possible choice of differentiable state variable we choose the canonical ones, the ones in which the density is expressed in appropriate units.
+Every requirement is physically motivated.
+No extra requirements are left.
+
 While the power of this shift can really be appreciated in the context of our broder work, it should be clear how with relative few starting points, elements of relativity (e.g.
@@
 And note how conservation of entropy is related to deterministic and reversible evolution.
-Also note how deterministic and reversible evolution means the that the system is only influenced by itself, which means it is not influenced by anything else, which means it is isolated, which indeed means energy is conserved.
-This surely this cannot be all coincidence.
+Also note how deterministic and reversible evolution means the that the system is only influenced by itself, which means it is not influenced by anything else, which means it is isolated, which it's linked to energy conservation.\footnote{To be precise, conservation of energy is not an invariant property.
+The fact that a Hamiltonian exists, though, is an invariant property.} This surely this cannot be all coincidence.
 
@@
 		\hline 
-		& Particle (classical) & an infinitesimal part of the system (i.e. the limit of recursive division) \\ 
+		& Classical particle & an infinitesimal part of the system (i.e. the limit of recursive division) \\ 
 		$\mathcal{S} =\mathsf{T}^*\mathcal{Q}$ & Phase space \newline (Tangent bundle) & the set of all possible states for particles upon which invariant distributions can be defined \\
@@
 
+	\bibitem{North} Jill North; The "Structure" of Physics: A Case Study, The Journal of Philosophy 106, no. 2 (2009), Pages 57-88, http://www.jstor.org/stable/20620153.
+
+	\bibitem{Curiel}Erik Curiel; Classical Mechanics Is Lagrangian; It Is Not Hamiltonian, The British Journal for the Philosophy of Science, Volume 65, Issue 2, 1 June 2014, Pages 269–321, https://doi.org/10.1093/bjps/axs034
+
+	\bibitem{Barrett1}Thomas William Barrett; On the Structure of Classical Mechanics, The British Journal for the Philosophy of Science, Volume 66, Issue 4, 1 December 2015, Pages 801–828, https://doi.org/10.1093/bjps/axu005
+
+	\bibitem{Barrett2}Thomas William Barrett; Equivalent and Inequivalent Formulations of Classical Mechanics, The British Journal for the Philosophy of Science, , axy017, https://doi.org/10.1093/bjps/axy017
+
+	\bibitem{carc1} G. Carcassi, C. A. Aidala: \emph{Assumptions of Physics} (in preparation). http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.
+
+	\bibitem{carc2} G.
+Carcassi, C.
+A.
+Aidala, D.
+J.
+Baker and L.
+Bieri; From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics, Journal of Physics Communications, 2, 4, 045026, 2018.
+
 \end{thebibliography}

### 2018-10-28 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Added Christine's comments

@@
 \title{Hamiltonian mechanics is conservation of information entropy}
-\author{Christine A. Aidala, Gabriele Carcassi}
+\author{Gabriele Carcassi, Christine A. Aidala}
 
@@
 	In this work we show the equivalence between Hamiltonian mechanics and conservation of information entropy.
-We will show that distribution with coordinate independent values for information entropy require that the manifold on which the distribution is defined is charted by conjugate pairs (i.e.
+We will show that distributions with coordinate independent values for information entropy require that the manifold on which the distribution is defined is charted by conjugate pairs (i.e.
 it is a symplectic manifold).
 We will also show that further requiring that the information entropy is conserved during the evolution yields Hamilton's equations.
-This will provide new insights in the nature of classical mechanics and physics in general, as it will connect to elements of relativity, statistical mechanics, thermodynamics, quantum mechanics, probability theory and symplectic geometry.
-This work is part of a larger project, called Assumptions of physics, which aims to rederive the known equations from a handful of physically meaningful starting points.
+This will provide new insights into the nature of classical mechanics and physics in general, as it will connect to elements of relativity, statistical mechanics, thermodynamics, quantum mechanics, probability theory and symplectic geometry.
+This work is part of a larger project, called Assumptions of Physics, which aims to rederive the known equations of physics from a handful of physically meaningful starting points.
 \end{abstract}
@@
 Over the past decade there has been a renewed interest in the underpinning of classical mechanics~\cite{North,Curiel,Barrett1,Barrett2}.
-Different attempts and arguments have been put forward to understand what are the most important feature of the different mathematical formulations, whether one is more fundamental of the other in some sense or whether they are equivalent.
+Different attempts and arguments have been put forward to understand what are the most important features of the different mathematical formulations, whether one is more fundamental than the other in some sense or whether they are equivalent.
 
@@
 Newtonian mechanics, thermodynamics and special relativity serve as good examples.
-Nowadays, unfortunately, much of the current fundamental theories, including Lagrangian, Hamiltonian, Quantum mechanics and Quantum Field Theory take as their starting point a particular mathematical structure.
-That is, we do not know what the value of a Lagrangian is or what the symplectic form represents: we just take them as a given.
+Nowadays, unfortunately, many of the current fundamental theories, including Lagrangian, Hamiltonian, quantum mechanics and quantum field theory take as their starting point a particular mathematical structure.
+That is, we do not know what the value of a Lagrangian means or what the symplectic form represents: we just take them as given.
 Moreover, a lot of work in theoretical physics nowadays starts by exploring novel abstract mathematical structures and then trying to see whether some physical meaning or prediction can be extracted.
 We are not going to debate here how and why this shift happened.
-We just merely note that it leaves the semantics of our theories ill defined.\footnote{Which is a nice way to say that we don't really know what are we talking about.}
+We merely note that it leaves the semantics of our theories ill defined.\footnote{Which is a nice way to say that we don't really know what we are talking about.}
 
@@
 The definitions of topology, differentiable manifold, symplectic manifold, Hilbert space and so on are chosen either because they are convenient to prove theorems or because they are convenient to make calculations.\footnote{In fact, it is common in math to have two such definitions and then prove they are equivalent.} We should not expect them to be a good match for definitions that would be physically meaningful.
-In other words, there is a very good chance that just looking at just the math may lead us, like horses with blinders, to a very clear path which is not the physically meaningful one.\footnote{To be clear, we do not fault the mathematicians either as they are not responsible for whether their structures are used appropriately or not.}
+In other words, there is a very good chance that looking at just the math may lead us, like horses with blinders, to a very clear path which is not the physically meaningful one.\footnote{To be clear, we do not fault the mathematicians either as they are not responsible for whether their structures are used appropriately or not.}
 
-Fully conscious of these issues, we have embarked in an effort, called Assumptions of Physics~\cite{carc1}, that aims to find a handful of physical principles from which the known laws of physics can be formally derived~\cite{carc2}.
+Fully conscious of these issues, we have embarked nn an effort, called Assumptions of Physics~\cite{carc1}, that aims to find a handful of physical principles from which the known laws of physics can be formally derived~\cite{carc2}.
 The approach has several advantages.
-First, by starting from physical ideas we put physics back at the center of the discussion, and not relegate it as a mere after the fact interpretation: the semantics is always well defined.
+First, by starting from physical ideas we put physics back at the center of the discussion, instead of relegating it to a mere after-the-fact interpretation: the semantics is always well defined.
 Second, it puts physics on sturdier mathematical grounds as we have a precise physical meaning to all the formal objects and we leave out those that do not.
-Third, it fosters connections between different fields of science, math, philosophy and engineering: it is our belief that nature is one and does not care about such divisions and therefore truly fundamental concepts must span freely across different areas.
-Fourth, it gives a framework to pose deep questions of the type the philosophy of science community is interested in.
-And lastly, by being forced to formalize and make explicit our assumptions, we can better understand how and why they fail possibly giving new insights that may lead to new physics.
-While the project is ambitious, we have already made progress beyond what we thought it could be reasonably done.
-So there's that.
+Third, it fosters connections between different fields of science, math, philosophy and engineering: nature is one and does not care about such divisions and therefore truly fundamental concepts must span freely across different areas.
+Fourth, it gives a framework to pose questions of the type in which the philosophy of science community is interested.
+And lastly, by being forced to formalize and make explicit our assumptions, we can better understand how and why they fail, possibly giving new insights that may lead to new physics.
+While the project is ambitious, we have already made progress beyond what we thought could be reasonably done.
 
@@
 a symplectic structure where variables are organized in conjugate pairs).
-In section 2 we show that, in general, information entropy and distributions densities are coordinate dependent.
+In section 2 we show that, in general, information entropy and densities are coordinate dependent.
 In section 3 we show that the structure of phase space is required to make such objects invariant under coordinate transformation (i.e.
 observer independent) and therefore give them an objective physical character.
-Conservation of entropy over time will force that invariance during time evolution, giving rise to Hamilton's equations.
-In chapter 4 we will briefly discuss the result and show how it necessarily leads to a classical version of the uncertainty principle.
+Conservation of entropy over time will enforce that invariance during time evolution, giving rise to Hamilton's equations.
+In section 4 we will briefly discuss the result and show how it necessarily leads to a classical version of the uncertainty principle.
 
@@
 This is the dual of determinism, that is the ability, given a state, to predict its future.} evolution of infinitesimally reducible objects}.
-As the information needed to identify an element of the distribution does not change in time, knowing the initial state is equivalent to knowing the future or past state.
+As the information needed to identify an element of the distribution does not change in time, knowing the current state is equivalent to knowing the future or past state.
 Classical particles, then, should be understood as the limit of subsequent subdivisions, infinitesimally small objects with extent in phase space.
-This single physical model is captured piecemeal by elements of statistics, probability theory and symplectic geometry which are given a coherent and well defined meaning.
+This single physical model is captured piecemeal by elements of statistics, probability theory and symplectic geometry which are now given a coherent and well defined meaning.
 In our view, this understanding also meshes better with other ideas from general relativity\footnote{A point particle in general relativity would not follow a geodesic and would, due to the infinite mass density, puncture space-time.}, thermodynamics and quantum mechanics.
@@
 
-While the mathematical setting is the same for degrees of belief (credence) or the likelyhood that something will happen (probability), we will assume we are working with an actual physical distribution.
+While the mathematical setting is the same for degrees of belief (credence) or the likelihood that something will happen (probability), we will assume we are working with an actual physical distribution.
 Even when we are talking about marginal distributions or information entropy, we are doing it in the context of a system that has parts and each part has a different characterization.
@@
 In general, if $I$ is the number of questions then $C = 2 ^ I$ are the number of cases you can distinguish with those questions.
-On the other hand, suppose $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
+On the other hand, suppose for a particular value $\xi_0$ of a discrete distribution we have $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
 That is, the normalized distribution is the inverse of the number of cases: $\rho(\xi_0) = \frac{1}{C}$.
-Therefore the information needed to identify a particular element is $I(\xi^a)=\log \frac{1}{\rho(\xi^a)}$.
+Therefore the information needed to identify a particular case is $I(\xi^a)=\log \frac{1}{\rho(\xi^a)}$.
 If we take the average and change the base of the logarithm, we have:
 \begin{equation}
-I(\rho(\xi^a)) = \int \rho(\xi^a)) \ln \left(\frac{1}{\rho(\xi^a))}\right) d\xi^1 d\xi^2 ... d\xi^n =-\int \rho(\xi^a)) \ln (\rho(\xi^a)) d\xi^1 d\xi^2 ... d\xi^n
+I(\rho(\xi^a)) = \int \rho(\xi^a) \ln \left(\frac{1}{\rho(\xi^a)}\right) d\xi^1 d\xi^2 ... d\xi^n =-\int \rho(\xi^a) \ln (\rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n
 \end{equation}
 which is the information entropy of the distribution.\footnote{The nature of information entropy is often misunderstood, especially in physics.
-It does not represent the information the distribution has: it represent the \emph{additional} information that would be needed to identify an element.
-Also, it does not represent knowledge: as we are studying actual physical distributions, any correct distribution has full knowledge of the system.}
+It does not represents the information the distribution has: it represent the \emph{additional} information that would be needed to identify an element.
+Also, it does not represent knowledge: as we are studying actual physical distributions, any correct distribution represents full knowledge of the system.}
 
 Technically, given the more convenient choice of the natural logarithm, the unit of information is the nat, which is equal to $\frac{1}{\ln 2}$ bits.
-However, in the discussion we will refer to it as measured in bits, as we believe that the majority of the readers is more conformable with the latter unit.
+However, in the discussion we will refer to it as measured in bits, as we believe that the majority of the readers is more comfortable with the latter unit.
 Note that for continuous variables the expression has a slightly different meaning: the uniform distribution in a unit volume is assigned zero entropy so $I(\rho(\xi^a))$ represents the relative number of bits to achieve the same level of precision, which can be negative.
@@
 There is one feature about densities we need to be fully aware of: they are not invariant under change of variables.
-Intuitively, if we change units of the denominator the distribution changes value.
+Intuitively, if we change units the density changes value.
 For example, $1$ kg/m becomes $1000$ kg/km: the number changes and the units allow us to understand that the overall quantity remains the same but mathematics does not keep track of units.
@@
 the density is multiplied by the absolute value of the Jacobian determinant of the transformation.
-As the Jacobian is in general a function of the variables themselves, this also means that whether the density at one point is greater or smaller than at another also depends on the choice of variables.
+As the Jacobian is in general a function of the variables themselves, this means that whether the density at one point is greater or smaller than at another also depends on the choice of variables.
 
@@
 The reason is that we may be changing scale and therefore level of precision.
-Given that both change is determined by the Jacobian, we conclude that \textbf{during a change of variable the information entropy of all possible distributions does not change if and only if the density remains the same at every point under the same change of variable}.
+Given that the change in both density and entropy is determined by the Jacobian, we conclude that \textbf{during a change of variable the information entropy of all possible distributions does not change if and only if the density remains the same at every point under the same change of variable}.
 
@@
 For each random variable we can define its cumulative distribution function $F_\xi(\xi_0)=\mu(\xi<\xi_0)$ and the probability density function is its Radon-–Nikodym derivative $\rho(\xi) = \frac{dF_\xi}{d\xi}$.
-That is: the integral is the primal object while the density is derived.
+That is, the integral is the primal object while the density is derived.
 
-This approach actually makes more physical sense: what we measure is the amount of mass in a finite region and the density is the limit for a smaller and smaller ones.
+This approach actually makes more physical sense: what we measure is the amount of mass in a finite region and the density is the limit for a smaller and smaller one.
 The fundamental objects (outcomes, events and probability) are all defined without reference to variables.
@@
 Then we define $n$-forms as multi-linear functions of $n$ vectors that return the value associated with the infinitesimal parallelepiped they form.
-That is $\nu : (\mathsf{T}^*\mathcal{M})^n \rightarrow \mathbb{R}$.
+That is $\nu : (\mathsf{T}\mathcal{M})^n \rightarrow \mathbb{R}$.
 As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $\mu(V)=\int_V \nu(dV)$ with no reference to coordinates.
-If $e_a$ are the basis vector associated with $\xi^a$ and $e^b$ are linear functions such that $e^b(e_a)=\delta_a^b$, then we can express each $n$-form as $\rho(\xi^a)e_1\wedge e_2 \wedge ...
+If $e_a$ are the basis vectors associated with $\xi^a$ and $e^b$ are linear functions such that $e^b(e_a)=\delta_a^b$, then we can express each $n$-form as $\rho(\xi^a)e_1\wedge e_2 \wedge ...
 \wedge e_n$.
-The density $\rho(\xi^a)$ is the component of the $n$-form expressed in the $\xi^a$ coordinates in the a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value.
-But negative determinants means changing orientation of the region of integration, so fixing that would bring another sign change and $\rho(\xi^a)$ would not change.
+The density $\rho(\xi^a)$ is the component of the $n$-form expressed in the $\xi^a$ coordinates in a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value.
+But a negative determinant means changing orientation of the region of integration, so fixing that would bring another sign change and $\rho(\xi^a)$ would not change.
 This is equivalent to saying that some coordinate systems are right handed and some are left handed.}
@@
 This approach gives us a way to understand which objects are coordinate independent and which aren't.
-But it has two issues: it does not allow to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.\footnote{Typically, a vector is defined as a map between a scalar function of the manifold to another scalar function to the manifold.
-Or as an equivalence class of trajectories.
+But it has two issues: it does not allow us to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.\footnote{Typically, a vector is defined as a map between a scalar function of the manifold to another scalar function of the manifold, or as an equivalence class of trajectories.
 It is a stretch to think of fluid velocities or electric fields in those terms.
@@
 
-So the situation is that we have a single clear physical objects we want to study, but different and somewhat disconnected mathematical frameworks to study it.
+So the situation is that we have a single clear physical object we want to study, but different and somewhat disconnected mathematical frameworks to study it.
 This should not be surprising as the frameworks are defined by mathematicians, as we said before, to solve their own needs.
@@
 
-Suppose we have a system which we can think of made of infinitesimally small parts, which we call particles.
+Suppose we have a system which we can think of as made of infinitesimally small parts, which we call particles.
 Suppose  that each infinitesimal part can be assigned a state from a state space $\mathcal{S}$.
@@
 How can we know if the density stays the same if we can't even compare the value of the densities at two different elements of the distribution?
-How can we tell whether the information entropy is conserved if we can't event define a unique value at a given time?
+How can we tell whether the information entropy is conserved if we can't even define a unique value at a given time?
 
-This leads us to the following insight: \textbf{if we want to define deterministc and reversible evolution over distributions, they must be invariant distributions.
-That is, the densities and information entropy must be independent on the choice of coordinates.} As we noted before, requiring one to be coordinate invariant will automatically require the other to be invariant as well.
+This leads us to the following insight: \textbf{if we want to define deterministic and reversible evolution over distributions, they must be invariant distributions.
+That is, the densities and information entropy must be independent of the choice of coordinates.} As we noted before, requiring one to be coordinate invariant will automatically require the other to be invariant as well.
 
@@
 Because the densities and the information entropy vary with the Jacobian, we need something more.
-Our state space $\mathcal{S}$ must allow to write densities whose value does not change under coordinate transformation.
+Our state space $\mathcal{S}$ must allow us to write densities whose value does not change under coordinate transformation.
 This seems to be impossible given what we saw before: how can it work?
@@
 If we change unit $\hat{q}=\hat{q}(q)$ we have two constraints: the one unit change and the unitary Jacobian, and these must be enough to determine how the vector components change.
-The components must be equal to the constraints and therefore they are exactly two variable $\xi^a = (q, k)$.
+The number of components must be equal to the number of constraints and therefore there are exactly two variables $\xi^a = (q, k)$.
 Then $k$ is the conjungate variable of $q$.
@@
 \dfrac{\partial \hat{q}}{\partial q} & \dfrac{\partial \hat{q}}{\partial p} \\[2.2ex]
-\dfrac{\partial \hat{p}}{\partial q} & \dfrac{\partial \hat{p}}{\partial p} \end{matrix} \right| = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial x} \frac{\partial \hat{q}}{\partial p} &= \{\hat{q}, \hat{p}\}
+\dfrac{\partial \hat{p}}{\partial q} & \dfrac{\partial \hat{p}}{\partial p} \end{matrix} \right| = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial q} \frac{\partial \hat{q}}{\partial p} &= \{\hat{q}, \hat{p}\}
 \end{aligned}
 \end{equation}
-which we recognize it to be the Poisson bracket. If the change of variable is simply a change of coordinates, a change of units, we have:
+which we recognize to be the Poisson bracket. If the change of variable is simply a change of coordinates, a change of units, we have:
 \begin{equation}
@@
 
-We can think of the evolution as a vector field on $\mathcal{S}$ of components $S = (S_q, S_q) = (\frac{dq}{dt}, \frac{dp}{dt})$ that gives the direction in which states move in time.
+We can think of the evolution as a vector field on $\mathcal{S}$ of components $S = (S_q, S_p) = (\frac{dq}{dt}, \frac{dp}{dt})$ that gives the direction in which states move in time.
 The Jacobian for time evolution will be
@@
 \begin{aligned}
-S = - curl(H) &= (\frac{\partial H}{\partial q}, - \frac{\partial H}{\partial p}) \\
+S &= \left(\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q}\right) \\
 \frac{dq}{dt} &= \frac{\partial H}{\partial p}  \\
@@
 and we recognize Hamilton's equations.
-Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it's not just that the density and the areas are conserved in time: is that they are the same for all canonical coordinates.
-That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is the exactly the evolution of (classical) distributions that preserve information entropy and densities}.
+Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it's not just that the density and the areas are conserved in time: it is that they are the same for all canonical coordinates.
+That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is precisely the evolution of (classical) distributions that preserve information entropy and densities}.
 
-We'd like to stress how, once the simple premise of invariant distributions is accepted, all elements of Hamiltonian mechanics are necessary and have a straight forward physical meaning.
-Not only that: because of the equivalent we know that everything in Hamiltonian mechanics can be understood with just that simple premise.
-And the premise is not about cotangent bundles of Lie group of quasi-q spaces on an infinite-dimensonal lattice.
+We'd like to stress how, once the simple premise of invariant distributions is accepted, all elements of Hamiltonian mechanics are necessary and have a straightforward physical meaning.
+Not only that: because of the equivalence we know that everything in Hamiltonian mechanics can be understood with just that simple premise.
+And the premise is not about cotangent bundles of Lie groups of quasi-q spaces on an infinite-dimensional lattice.
 It's just the simple realization that if we require distributions to have a physically objective reality they must be coordinate independent.
@@
 
-To understand what happens in multiple degrees of freedom, let's first look at change of variables of marginal distributions.
+To understand what happens with multiple degrees of freedom, let's first look at change of variables of marginal distributions.
 Suppose we have a distribution $\rho(\xi^a)$ that depends on many coordinates, on many units.
@@
 where $\xi^b$ are the remaining coordinates.
-If the variables $\xi^b$ are independent from $\xi^1$ and $\xi^2$, then we must be able to change units independently as well.
+If the variables $\xi^b$ are truly independent from $\xi^1$ and $\xi^2$, then we must be able to change units independently as well.
 Which means $\rho(\xi^b)$ must be coordinate invariant.
 
-We can imagine to proceed iteratively, one unit at a time.
-We'll find that our variables are paired $(q^i, p_i)$  with the first defining the units and the seconds defining the invariant areas.
-We call each pair an independent degree of freedom as it constitute an independent choice of unit.
+We can imagine proceeding iteratively, one unit at a time.
+We'll find that our variables are paired $(q^i, p_i)$  with the first ones defining the units and the second ones defining the invariant areas.
+We call each pair an independent degree of freedom as it constitutes an independent choice of unit.
 We recognize $\mathcal{Q}$, the manifold charted by $q^i$, as the (poorly named) configuration space.
@@
 The state space for the particles, then, is a set of coordinates plus an equal number of conjugate variables that vary like covector components, therefore we recognize $\mathcal{S}=\mathsf{T}^*\mathcal{Q}$ as the cotangent bundle.\footnote{This does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number.
-The state space is just a manifold and the structure it has comes from how units transform, not because the variable represent different objects.
+The state space is just a manifold and the structure it has comes from how units transform, not because the variables represent different objects.
 Unfortunately mathematical structures strip out the units, so that knowledge is lost and then resurfaces in other structures.
-Changing state variables from position and momentum to position and velocity is not, as one may be mistakenly understand from the math, a change of objects: is relabeling the same states much like changing spatial coordinates relabels positions.
+Changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects: it is relabeling the same states much like changing spatial coordinates relabels positions.
 It is not, in general, a canonical transformation as the density will not correspond to the invariant one, and it is not simply a change of coordinates.
@@
 This means that relationships between the entropy of the total and marginal distributions are preserved.
-For example, saying that the marginal distributions $\rho_2(q^2)$ and $\rho_1(q^1)$ are independent is equivalent to saying that the entropy of the joint distribution is equal to the sum of the entropy of the marginals: $I(\rho(q^1,q^2)) = I(\rho_2(q^2)) +I(\rho_1(q^1))$.\footnote{This is also equivalent to saying that the joint distribution is the product of the marginals $\rho(q^1,q^2) = \rho_2(q^2)\rho_1(q^1)$.} That is, giving both values is equivalent to giving one and then the other.
+For example, saying that the marginal distributions $\rho_1(q^1, p_1)$ and $\rho_2(q^2, p_2)$ are independent is equivalent to saying that the entropy of the joint distribution is equal to the sum of the entropy of the marginals: $I(\rho(q^1,q^2, p_1, p_2)) = I(\rho_1(q^1, p_1)) +I(\rho_2(q^2, p_2))$.\footnote{This is also equivalent to saying that the joint distribution is the product of the marginals $\rho(q^1,q^2, p_1, p_2) = \rho_1(q^1, p_1)\rho_2(q^2, p_2)$.} That is, giving both values is equivalent to giving one and then the other.
 This means we can think of these relationships and quantities as real physical properties assigned to the degree of freedom and not just to the coordinates.
@@
 is zero if $\Sigma$ is a closed surface.
-To understand why, note that it corresponds to the case where the density is the unit constant.
-If the surface is a parallelopiped in some variables, each face that give a positive contribution will have a parallel and equal face that would give an opposite contribution.
+To understand why, note that it corresponds to the case where $\rho(\xi^a)=1$.
+If the surface is a parallelopiped in some variables, each face that gives a positive contribution will have a parallel face that would give an equal and opposite contribution.
 As the closed integral needs to be zero for all parallelopipeds, it will also be zero for an arbitrary closed surface.
@@
 That is, phase space is a symplectic manifold.
-And, as any symplectic manifald, we can find Darboux variables\footnote{In general, this gives a general understanding of what a differentiable structure is.
-Any manifold, even a non-differentiable one, can identify points with coordinates.
-A differentiable structure on a manifold selects a set of charts that allow differentiable coordinate change.
-But there are many such sets, which one do we pick?
-We pick the one that allows us to write densities for our distributions.
-Similarly, a symplectic manifolds selects only the coordinates that allows us to write invariant densities.
-Also note how both the tangent space and the space of distributions are vector spaces.
-In a way, the linearity of distributions corresponds to the linearity of the tangent space.} $q^i, p_j$ and express the two-form as
+And, as any symplectic manifald, we can find Darboux variables $q^i, p_j$ and express the two-form as
 \begin{equation}
@@
 As $\omega$ is preserved during coordinate transformations, the area within each degree of freedom is preserved meaning we are preserving the marginal distribution as well.
-Across degree of freedom, instead, we are preserving the angle meaning two perpendicular degrees of freedom remain perpendicular.
-But here perpedicular means independent: the density within the four dimensional volume is the product of the densities on sides.
+Across degrees of freedom, instead, we are preserving the angle meaning two perpendicular degrees of freedom remain perpendicular.
+But here perpedicular means independent: the density within the four dimensional volume is the product of the densities on the sides.
 As these are preserved under coordinate transformation, these now can correspond to actual physical entities.
@@
 Deterministic and reversible evolution will preserve the densities and therefore preserve $\omega$.
-That is, $\omega(v, w) = \omega(v',w')$ where $v'$ and $w'$ are the vector evolved in time.
-In fancy math terms, deterministic and reversible evolution for a distribution is a symplectomorphism.
+That is, $\omega(v, w) = \omega(v',w')$ where $v'$ and $w'$ are the vectors evolved in time.
+In math terms, deterministic and reversible evolution for a distribution is a symplectomorphism.
 As before, $S = \left(\frac{d\xi^a}{dt}\right)$ is the vector field corresponding to the evolution.
-The vector components change according to $v'^a = \partial_b \xi(t+dt)^a v^b = (\delta^a_b + \partial_b S^a dt) v^b$.
+The vector components change according to $v'^a = \partial_b \xi^a(t+dt) v^b = (\delta^a_b + \partial_b S^a dt) v^b$.
 Therefore we have:
@@
 
-The symplectic structure was recovered by requiring invariant distributions which in turns requires invariant integration over independent degrees of freedom.
+The symplectic structure was recovered by requiring invariant distributions which in turn requires invariant integration over independent degrees of freedom.
 Note how the requirement was given outside of the symplectic structure itself and therefore it cannot be recovered simply from that.
@@
 Then, if we are so keen, put distributions on it which turn out, somewhat magically, to have density and entropy defined in a coordinate independent way.
-It makes a lot more sense to start by saying we have a finite system which we pretend to be made of infinitesimal parts, which we keep to dividing as we please and derive the state space of particles as the limit of such process of subdivision.
-This gives uniquely gives us the actual mathematical structure and its physical meaning.
+It makes a lot more sense to start by saying we have a finite system which we pretend to be made of infinitesimal parts, which we keep dividing as we please and derive the state space of particles as the limit of such process of subdivision.
+This simultaneously gives the actual mathematical structure and its physical meaning.
 
-Not only the second approach is preferable because with fewer starting points it justifies the state space, the evolution and why that state space is so compatible with statistical mechanics.
-It also maps better to what we do in classical physics: we study composite systems (planets, cannonballs, beads on a wire, fluids) which we sometime pretend we can recursively divide.
+Not only is the second approach preferable because with fewer starting points it justifies the state space, the evolution and why that state space is so compatible with statistical mechanics, it also maps better to what we do in classical physics.
+We study composite systems (planets, cannonballs, beads on a wire, fluids) which we sometime pretend we can recursively divide.
 None of those objects, in fact, has a single position or momentum: only the center of mass has a single position and momentum but those are precisely the expectation values $\int q \rho(q,p) dq dp$ and $\int p \rho(q,p) dq dp$ over the distribution.
 In other words, particles are not point-like in classical mechanics.
-They are infinitesimal cells of phase-space and that's why they evolution is symplectomorphism: that's why phase space requires a differentiable structure.
+They are infinitesimal cells of phase-space and that's why their evolution is symplectomorphism.
 
@@
 If one is convinced that that is not realistic then he is already convinced that classical mechanics will not hold in those regimes.
-And if that simplification fails, it's not that we need an altogether new set of rules: the macroscopic object is still the same.
+And if that approximation fails, it's not that we need an altogether new set of rules: the macroscopic object is still the same.
 We simply need to change how the subdivision works.
@@
 The state space of particles is a manifold because we use real numbers to identify the states.
-Among all possible choice of state variables, we conveniently choose the ones in which the distribution can be expressed as a density.
+Among all possible choices of state variables, we conveniently choose the ones in which the distribution can be expressed as a density.
 That's the differentiable structure on the manifold.
@@
 Moreover the differentiable manifold has a symplectic structure because the density is invariant under coordinate transformation.
-Therefore among all possible choice of differentiable state variable we choose the canonical ones, the ones in which the density is expressed in appropriate units.
+Therefore among all possible choices of differentiable state variables we choose the canonical ones, the ones in which the density is expressed in appropriate units.
 Every requirement is physically motivated.
@@
 
-While the power of this shift can really be appreciated in the context of our broder work, it should be clear how with relative few starting points, elements of relativity (e.g.
+While the power of this shift can really be appreciated in the context of our broader work, it should be clear how with relative few starting points, elements of relativity (e.g.
 coordinate invariance), statistical mechanics (e.g.
@@
 And note how conservation of entropy is related to deterministic and reversible evolution.
-Also note how deterministic and reversible evolution means the that the system is only influenced by itself, which means it is not influenced by anything else, which means it is isolated, which it's linked to energy conservation.\footnote{To be precise, conservation of energy is not an invariant property.
-The fact that a Hamiltonian exists, though, is an invariant property.} This surely this cannot be all coincidence.
+Also note how deterministic and reversible evolution means that the system is only influenced by itself, which means it is not influenced by anything else, which means it is isolated, which is linked to energy conservation.\footnote{To be precise, conservation of energy is not an invariant property.
+The fact that a Hamiltonian exists, though, is an invariant property.} Surely this cannot be all coincidence.
 
@@
 
-Having a coherent picture across the different areas not only is intellectually more satisfying, but it allows us to switch perspective when the intuition from one fails.
+Having a coherent picture across the different areas is not only intellectually more satisfying, but it allows us to switch perspective when the intuition from one fails.
 As a practical example, we can show how understanding the link between Hamiltonian mechanics and information theory leads very straightforwardly to a classical equivalent for the uncertainty principle.
@@
 \end{align*}
-Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and their elements, and therefore it can't shrink it more than a set amount.
+Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and its elements, and therefore it can't shrink the distribution more than a set amount.
 If it did, it would start concentrating the density, meaning that multiple states have part of their density mapped to a single state and the dynamics would not be reversible.
@@
 
-In the narrower context, it allows to understand what physical object is represented by each mathematical one. This can be summarized in table \ref{dictionary}.
+In the narrower context, it allows us to understand what physical object is represented by each mathematical one. This can be summarized in table \ref{dictionary}.
 \begin{table}[h]
@@
 		& Classical particle & an infinitesimal part of the system (i.e. the limit of recursive division) \\ 
-		$\mathcal{S} =\mathsf{T}^*\mathcal{Q}$ & Phase space \newline (Tangent bundle) & the set of all possible states for particles upon which invariant distributions can be defined \\
+		$\mathcal{S} =\mathsf{T}^*\mathcal{Q}$ & Phase space \newline (Cotangent bundle) & the set of all possible states for particles upon which invariant distributions can be defined \\
 		$\xi^a$ & State variables \newline (Unified coordinates) & the set of variables needed to identify the state of a particle \\ 
@@
 		$p_i=\hbar k_i$ & Conjugate momentum & a state variable that together with the corresponding coordinate defines invariant ranges of possible states \\
-		$(q^i, p_i)$ & Canonical variables \newline (Canonical coordinates) & a set of variable for which the density is invariant and expressed over units of $\hbar$\\ 
+		$(q^i, p_i)$ & Canonical variables \newline (Canonical coordinates) & a set of variables for which the density is invariant and expressed over units of $\hbar$ per degree of freedom\\ 
 		& Canonical transformation & a change in state variables that does not change the density\\ 
@@
 \end{table}
-It also allows us to see that deterministic and reversible evolution must conserve entropy and that isolated systems must conserve energy.
-In the broader context, it allows to make deeper connections between different areas of math and physics, and pushes us to rethink our starting points.
+It also allows us to see why deterministic and reversible evolution must conserve entropy and why isolated systems must conserve energy.
+In the broader context, it allows us to make deeper connections among different areas of math and physics, and pushes us to rethink our starting points.
 In particular, it tells us that classical particles should not be considered point-like objects (upon which entropy and densities cannot be defined), but the limit of recursive subdivisions (i.e.
@@
 
-These insights come from our project, Assumptions of Physics, that aims to re-derive the standard laws from a handful of physical starting points.
+These results represent one set of insights from our larger project, Assumptions of Physics, that aims to rederive the standard laws from a handful of physical starting points.
 We believe that a reorganization of fundamental physics on more physically meaningful conceptual footing and more rigorous mathematical grounds is not only possible, but long overdue.
-The insights presented here are one step toward that direction as we know them to be still valid and useful in a larger setting, such as understanding the similarities in the mathematical structures of classical and quantum mechanics.
 

### 2018-11-17 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Reworking first two section based on Mario's feedback

@@
 We will also show that further requiring that the information entropy is conserved during the evolution yields Hamilton's equations.
-This will provide new insights into the nature of classical mechanics and physics in general, as it will connect to elements of relativity, statistical mechanics, thermodynamics, quantum mechanics, probability theory and symplectic geometry.
-This work is part of a larger project, called Assumptions of Physics, which aims to rederive the known equations of physics from a handful of physically meaningful starting points.
 \end{abstract}
@@
 
-Over the past decade there has been a renewed interest in the underpinning of classical mechanics~\cite{North,Curiel,Barrett1,Barrett2}.
-Different attempts and arguments have been put forward to understand what are the most important features of the different mathematical formulations, whether one is more fundamental than the other in some sense or whether they are equivalent.
+Over the past decade there has been a renewed interest in the underpinning of classical mechanics.
+Different attempts and arguments have been put forward to understand what are the most important features of the different mathematical formulations, whether they are equivalent or whether one is more fundamental than the other~\cite{North,Curiel,Barrett1,Barrett2}.
+We believe, though, that there are two general problems with these attempts.
 
-All these attempts, though, suffer from one problem.
-They concentrate on the mathematical structure.
-The issue is that, in general, a mathematical object can represent more than one physical entity.
-For example, the same mathematical framework for linear circuits can be used to describe electric, hydraulic, thermal or mechanical systems.
-On the other end, the same physical entity can be described by more than one mathematical object.
-For example, the same cannonball can be described by position and velocity in a Lagrangian setting, position and momentum in a Hamiltonian setting or in a thermodynamics setting by temperature, volume and so on.
-The connection between math and physics, then, is many-to-many and therefore we can't study only one side.
-The reason is that mathematics captures only the syntax of a particular description without its semantics, which at that point cannot be recovered.
-Mathematical structures are just shadows of the physical model they represent.
+The first is that they concentrate on the mathematical structure which, unfortunately, is not enough to characterize a physical theory.
+They all note how the same physical system can be described by different mathematical frameworks.
+For example, a classical system can be characterized by position and velocity in the Lagrangian framework and by position and momentum in the Hamiltonian framework.
+They do not note, though, that the converse is also true: the same mathematical equation can represent two different physical systems.
+For example, the equation $ma + bv = 0$ in the context of Newtonian mechanics can either represent a body in an inertial frame under linear drag or a body in a non-inertial frame with no forces acting on it.
+If $t_{in}$ is time in the inertial frame, the time $t$ in the non-inertial frame is such that  $\frac{dt}{dt_{in}} = e^{\frac{b}{m}t}$.
+Mathematical equivalence does not imply physical equivalence, which is especially problematic for \cite{Barrett2} since it focuses on the former.
+In other words: we can have the same physics with different math and the same math with different physics.
+The connection between math and physics is many-to-many because mathematics captures only the syntax of a particular description without its semantics, which at that point cannot be recovered.
+
+As we do not have a one-to-one map between math and physics, fixating on the mathematical structure may even confuse us.
+For example, it is known\cite{AllHamFreeParticle} that any Hamiltonian system is, at least locally, equivalent to a free particle.
+That is, we can always find a canonical transformation (i.e.
+a transformation that does not change the equation of motion) such that any Hamiltonian system, in the new coordinates, has the motion of a free particle for a finite amount of time.
+As \cite{North} notes, only frame independent relationships should be taken as fundamental.
+Does that mean that the Hamiltonian itself is not a fundamental relationship, since it is coordinate dependent and all Hamiltonians are the same anyway?
+It is also known\cite{AllSystemsAreHam} that for any first order system of ordinary differential equations we can find an appropriate symplectic structure such that the system can be written as a Hamiltonian system.
+Does that mean that, in the end, all systems are physically equivalent, regardless of what they describe?
+The point is, we can't know whether our findings about mathematical structures are physically significant if we do not have a precise link between the mathematical symbols and their physical meaning.
+
+This leads us to the second problem we believe the previous works may have: when trying to attach meaning the thinking and intuition is still Newtonian.
+This is particularly evident in \cite{Curiel} where we start from physical ideas to get to Lagrangian dynamics.
+The starting point is that a classical system is one that, roughly, obeys Newton's second law and therefore we have, for a free particle, $\dot{x} = v$ and $\dot{v} = 0$.
+Unfortunately this is only true in an inertial frame.
+As \cite{North} notes, both Hamiltonian and Lagrangian mechanics are coordinate independent: the equations are valid in all reference frames even non-inertial ones.
+In fact, that is probably the most notable advantage of the frameworks, that fictitious forces are automatically handled.
+But none of the Newtonian concepts are coordinate independent: inertial frames, absence of forces and even conservation of energy are all frame dependent.
+This, again, leads to confusion and we may miss a key element: in the transition from Newtonian to Lagrangian/Hamiltonian mechanics we expand the reference frames in which the equations are valid (from inertial frames to all frames) at the expense of restricting the class of systems we can describe (e.g.
+dissipative forces in general require a modification of the Euler-Lagrange equation using the Rayleigh dissipation function).
+The point is, Newtonian concepts may prevent us from fully understanding Hamiltonian and Lagrangian mechanics.
 
 It should be clear that we do not fault the authors.
-Even just a century ago, there was a push to find fundamental principles or laws that could serve as a starting point for the different theories.
+We fault us, the physicists.
+Even just a century ago, there was a push in our field to find fundamental principles or laws that could serve as a starting point for the different theories.
 Newtonian mechanics, thermodynamics and special relativity serve as good examples.
@@
 
-It is only natural that, given the circumstances, one uses the only thing that seems to be well defined, the math, and tries to get understanding from there.
+It is only natural that, given the circumstances, a philosopher would use the only thing that seems to be well defined, the math, and try to draw understanding from there.
 But here lies the fundamental problem: that math is not created by physicists to do physics, it is created by mathematicians to do math.
-The definitions of topology, differentiable manifold, symplectic manifold, Hilbert space and so on are chosen either because they are convenient to prove theorems or because they are convenient to make calculations.\footnote{In fact, it is common in math to have two such definitions and then prove they are equivalent.} We should not expect them to be a good match for definitions that would be physically meaningful.
-In other words, there is a very good chance that looking at just the math may lead us, like horses with blinders, to a very clear path which is not the physically meaningful one.\footnote{To be clear, we do not fault the mathematicians either as they are not responsible for whether their structures are used appropriately or not.}
+The definitions of topology, differentiable manifold, symplectic manifold, Hilbert space and so on are chosen either because they are convenient to prove theorems or because they are convenient to perform calculations.\footnote{In fact, it is common in math to have two such definitions and then prove they are equivalent.} We should not expect them to be a good match for definitions that would be physically meaningful.
+In other words, there is a very good chance that looking at just the math may lead us, like horses with blinders, in a well-defined direction which is not the physically meaningful one.
+To be clear, we do not fault the mathematicians either as they are not responsible for whether their structures are used appropriately or not within physics.
 
-Fully conscious of these issues, we have embarked nn an effort, called Assumptions of Physics~\cite{carc1}, that aims to find a handful of physical principles from which the known laws of physics can be formally derived~\cite{carc2}.
-The approach has several advantages.
-First, by starting from physical ideas we put physics back at the center of the discussion, instead of relegating it to a mere after-the-fact interpretation: the semantics is always well defined.
-Second, it puts physics on sturdier mathematical grounds as we have a precise physical meaning to all the formal objects and we leave out those that do not.
-Third, it fosters connections between different fields of science, math, philosophy and engineering: nature is one and does not care about such divisions and therefore truly fundamental concepts must span freely across different areas.
-Fourth, it gives a framework to pose questions of the type in which the philosophy of science community is interested.
-And lastly, by being forced to formalize and make explicit our assumptions, we can better understand how and why they fail, possibly giving new insights that may lead to new physics.
-While the project is ambitious, we have already made progress beyond what we thought could be reasonably done.
+As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work\cite{AoPPhy1} that aims to identify such starting points.
+Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
+a symplectic structure where variables are organized in conjugate pairs).
 
-In this paper we focus on just one particular insight provided by our work: \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
-a symplectic structure where variables are organized in conjugate pairs).
-In section 2 we show that, in general, information entropy and densities are coordinate dependent.
-In section 3 we show that the structure of phase space is required to make such objects invariant under coordinate transformation (i.e.
-observer independent) and therefore give them an objective physical character.
-Conservation of entropy over time will enforce that invariance during time evolution, giving rise to Hamilton's equations.
-In section 4 we will briefly discuss the result and show how it necessarily leads to a classical version of the uncertainty principle.
+We believe this insight may be of interest to the philosophy of science community for at least two reasons.
+First because it helps to clearly characterize what Hamiltonian mechanics describes, providing a clear physical meaning to the mathematical structures used.
+This should be at least of interest to that part of the philosophy community interested in the foundations of classical mechanics.
+Second because information entropy has a deep connection to the foundations of thermodynamics, therefore the strong link we establish here between information entropy and Hamiltonian mechanics should be of interest to an even wider audience.
 
-The overall idea is that \textbf{Hamiltonian mechanics is best understood as the deterministic and reversible\footnote{By reversible we mean the ability, given a state, to reconstruct its past.
-This is the dual of determinism, that is the ability, given a state, to predict its future.} evolution of infinitesimally reducible objects}.
-As the information needed to identify an element of the distribution does not change in time, knowing the current state is equivalent to knowing the future or past state.
-Classical particles, then, should be understood as the limit of subsequent subdivisions, infinitesimally small objects with extent in phase space.
-This single physical model is captured piecemeal by elements of statistics, probability theory and symplectic geometry which are now given a coherent and well defined meaning.
-In our view, this understanding also meshes better with other ideas from general relativity\footnote{A point particle in general relativity would not follow a geodesic and would, due to the infinite mass density, puncture space-time.}, thermodynamics and quantum mechanics.
+We begin in section 2 by giving a brief summary about statistical distributions in general.
+These are your typical mathematical tools used to keep track of how charge is distributed in space or how a population of individuals is distributed in age.
+We briefly introduce information entropy, which measures the number of yes or no questions one has to answer to be able to identify a particular element within a distribution.
+The main issue is that the density and the information entropy for a distribution depend, in general, on the variables and their units.
+For example, if we change our units from meters to kilometers a density of $1 kg/m^3$ would change to $10^9 kg/km^3$.
+In other words, distributions are not in general invariant under change of variables.
+In section 3 we look at distributions over states.
+We argue that since states themselves are coordinate independent, the density and information entropy associated with them cannot be coordinate dependent.
+Moreover, a deterministic and reversible system must be one that preserves information entropy over time (i.e.
+information about the system at one time is equivalent to information about the system before or after).
+The requirement of coordinate independent distributions gives us the symplectic structure (i.e.
+conjugate pairs of position and momentum) while the requirement of conservation of information entropy gives us Hamiltonian evolution.
+In section 4 we discuss the result and provide a dictionary that allows us to give precise meaning to the different mathematical symbols, addressing the issues we highlighted before.
+We also show how this understanding allows us to construct a classical analogue of the uncertainty principle.
 
@@
 
-The general idea is that we want to describe a quantity that is distributed among different cases.
-For example, mass or charge over different positions in space, amount of material among different compounds, population at different ages.
-In all these cases, we assume there is a total quantity $M$, a set of $n$ variables $\xi^a$ over which the quantity is distributed and a function $\rho(\xi^a)$ such that:
+Our starting point is the standard concept of a statistical distribution.
+We have a set of elements, also called population, and a set of variables that characterize some properties of those elements.
+To each element is associated a value for each variable and we want to describe how the population is distributed over the possible values of the variables.
+For example, people by age and country of residence, amount of material among different compounds, mass or charge over different positions in space.
+In all these cases, we have a total quantity $M$ that measures how many elements we have.
+This can be either a discrete quantity, like the number of people we are considering, or a continuous quantity, like the total amount of mass.
+We also have a set of $n$ variables $\xi^a$ over which the elements are distributed and a function $m(\xi^a)$ that gives us the amount for each value possible value such that:
+\begin{equation}
+\int m(\xi^a) d\xi^1 d\xi^2 ... d\xi^n = M
+\end{equation}
+The integral is used when some of the variables $\xi^a$ are continuous and a simple sum would be used over the discrete ones. We can also define the normalized density as:
+\begin{equation}
+\rho(\xi^a) = \frac{1}{M}m(\xi^a)
+\end{equation}
+To calculate the amount in a particular region $V$ we have:
+\begin{equation}
+m_V = \int_V M \rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n
+\end{equation}
+The above expression would represent the mass or charge within a particular region, the amount of material of some selected set of compounds or the world population in a particular set of countries and age group.
+We can also define:
+\begin{equation}
+\mu(V) =\frac{m_V}{M} = \int_V \rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n \\
+\end{equation}
+which gives the fraction of the population within that region.
 
-\begin{equation}
-\int M \rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n = M
-\end{equation}
+The normalized distribution $\rho(\xi^a)$, as defined, is a statistical distribution that represents the entirety of the system being described.
+Yet, we can give it an additional meaning of probability if we run the following thought experiment: suppose we took an element at random, each element with equal chance, what is the probability $P(\xi^a \in V)$ that the values for $\xi^a$ associated with that element are within a particular region $V$?
+The answer must be $\mu(V)$ and therefore $\rho(\xi^a)$ is the probability density.
+So $\rho(\xi^a)$ has this double role as a physical statistical distribution and a conceptual probability distribution for our thought experiment.
+This second role is our link to information theory.
 
-We say that $\rho$ is the (normalized) density for our distribution and if we want to calculate the amount in a particular region $V$ we have:
+Before proceeding, though, let's give a quick introduction so that we have a clear understanding about what information entropy represents.
+In the physics community, for example, the concept is often vaguely referred to as ``knowledge" (and sometimes as ``lack of knowledge") which is incorrect and misleading.\footnote{We believe this characterization may be due to Jaines\cite{Jaines} who first introduced the concept of information theory within thermodynamics.} Information entropy was introduced by Shannon\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and as such should be treated.
+What he needed was a way to measure how much information a particular channel is able to transfer and therefore he needed a way to quantify precisely information as number of bits.
+So the idea is that we have a source that sends messages to a destination, and the message is encoded in zeros and ones.
+Information entropy is the optimal average number of zeros and ones that must be used to encode the messages.
+Below, the message is crippled and above the message is redundant.
 
-\begin{equation}
-m_V = M \mu(V) =\int_V M \rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n
-\end{equation}
-
-The above expression would represent the mass or charge within a particular region, the amount of material of some selected set of compounds or the population in a particular age group.
-The integral is used when some of the variables $\xi^a$ are continuous and a simple sum would be used over the discrete ones.
-
-While the mathematical setting is the same for degrees of belief (credence) or the likelihood that something will happen (probability), we will assume we are working with an actual physical distribution.
-Even when we are talking about marginal distributions or information entropy, we are doing it in the context of a system that has parts and each part has a different characterization.
-
-The link to information theory is the following: suppose we have a distribution, known to both of us, and you select one element at random.
-How many yes or no questions do you have to answer before I know what case you selected?
-For example, suppose there are 4 balls of which 2 are red and 2 are green, one yes or no question would be sufficient.
-In general, if $I$ is the number of questions then $C = 2 ^ I$ are the number of cases you can distinguish with those questions.
-On the other hand, suppose for a particular value $\xi_0$ of a discrete distribution we have $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
+To understand how this work in our context, let's go back to our distribution $\rho(\xi^a)$ and refine our thought experiment.
+Suppose we have a distribution, known to both us.
+Suppose you pick an element at random and I need to know the value of the variable for the element you picked.
+That is, you are the source, I am the destination and the message is the value of the element you picked from the distribution.
+How long, on average, should the message be?
+How many ones and zeros?
+How many yes or no questions must I ask you before I know the value?
+For example, suppose there are 40 balls of which 20 are red and 20 are green.
+In that case one yes or no question (``Is it green?"), one bit of information, would be sufficient.
+Suppose there are 40 balls of which 10 red, 10 green, 10 blue and 10 yellow.
+In this case, two questions would suffice as two yes or no questions, two bits of information, can distinguish between 4 cases.
+In general, if $I$ is the number of questions, the number of bits, then $C = 2 ^ I$ are the number of cases you can distinguish with those questions.
+On the other hand, suppose for a particular value $\xi_0$ of a discrete variable we have $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
 That is, the normalized distribution is the inverse of the number of cases: $\rho(\xi_0) = \frac{1}{C}$.
@@
 \end{equation}
-which is the information entropy of the distribution.\footnote{The nature of information entropy is often misunderstood, especially in physics.
-It does not represents the information the distribution has: it represent the \emph{additional} information that would be needed to identify an element.
-Also, it does not represent knowledge: as we are studying actual physical distributions, any correct distribution represents full knowledge of the system.}
+which is the information entropy of the distribution.
+That tells us the average number of bits you need to put into a message to tell me the value of the element of the distribution you picked at random.
 
@@
 However, in the discussion we will refer to it as measured in bits, as we believe that the majority of the readers is more comfortable with the latter unit.
-Note that for continuous variables the expression has a slightly different meaning: the uniform distribution in a unit volume is assigned zero entropy so $I(\rho(\xi^a))$ represents the relative number of bits to achieve the same level of precision, which can be negative.
+Also, technically, that for continuous variables the expression has a slightly different meaning: the uniform distribution in a unit volume is assigned zero entropy so $I(\rho(\xi^a))$ represents the relative number of bits to achieve the same level of precision, which can be negative.
+That is, if we take a real number from a uniform distribution between 0 and $\frac{1}{2}$ we need to loose a bit of information to get to the same precision of a unitary distribution.
+
+Technicality aside, the above discussion, while not a formal derivation, should give a sense of what the expression signifies: it measures the \emph{additional} information that would be needed to identify the value associated to an element picked at random from a known population.
+What is also important is that it is a precisely defined quantity and that there is only one way to define it.
+It is the only expression that satisfies three simple requirements (continuity, monotonicity and additivity) that are necessary to measure information and that is how Shannon originally derived it.
+Insofar that we want to talk about amounts of information we have only one way to do it.
 
@@
 
-There is one feature about densities we need to be fully aware of: they are not invariant under change of variables.
+There is one feature about densities we need to be fully aware of: in general they are not invariant under change of variables.
 Intuitively, if we change units the density changes value.
@@
 \end{equation}
-The reason is that we may be changing scale and therefore level of precision.
+The reason is that we may be changing scale and therefore level of precision required to describe our value.
 Given that the change in both density and entropy is determined by the Jacobian, we conclude that \textbf{during a change of variable the information entropy of all possible distributions does not change if and only if the density remains the same at every point under the same change of variable}.
@@
 To sum up, while we'd like to think of a density as a number associated to a point of a space, that's technically incorrect.
-Given a manifold $\mathcal{M}$ upon which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \to \mathbb{R}$.
+Given a manifold $\mathcal{M}$ upon which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \to \mathbb{R}$ in general.
 That would be a scalar function, invariant under change of variables, as a point is a point no matter how is it identified.
@@
 
-This approach actually makes more physical sense: what we measure is the amount of mass in a finite region and the density is the limit for a smaller and smaller one.
+This approach, if applied to statistical distributions, actually makes more physical sense: what we measure is the amount of mass in a finite region and the density is the limit for a smaller and smaller one.
 The fundamental objects (outcomes, events and probability) are all defined without reference to variables.
-But in probability theory we have no notion of coordinate systems, vectors and all other geometrical notions.
+It also has the advantage that discrete and continuous quantities are treated in the same way.
+But in probability theory we have no notion of coordinate systems, vectors and all other geometrical objects.
 
@@
 That is $\nu : (\mathsf{T}\mathcal{M})^n \rightarrow \mathbb{R}$.
-As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $\mu(V)=\int_V \nu(dV)$ with no reference to coordinates.
+As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $\mu(V)=\int_V \nu$ with no reference to coordinates.
 If $e_a$ are the basis vectors associated with $\xi^a$ and $e^b$ are linear functions such that $e^b(e_a)=\delta_a^b$, then we can express each $n$-form as $\rho(\xi^a)e_1\wedge e_2 \wedge ...
@@
 This approach gives us a way to understand which objects are coordinate independent and which aren't.
-But it has two issues: it does not allow us to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.\footnote{Typically, a vector is defined as a map between a scalar function of the manifold to another scalar function of the manifold, or as an equivalence class of trajectories.
-It is a stretch to think of fluid velocities or electric fields in those terms.
-For that same reason, we do not use the differential geometry notation $\frac{\partial}{\partial \xi^a}$ and $d\xi^a$ for vector and covector basis because they do not convey the actual physical meaning of these objects as they are used in physics.}
+But it has two issues: it does not allow us to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.
+For example, a vector is typically defined as a map between a scalar function of the manifold to another scalar function of the manifold, or as an equivalence class of trajectories.
+It is a stretch to think of fluid velocities or electric fields in those terms.\footnote{For that same reason, we do not use the differential geometry notation $\frac{\partial}{\partial \xi^a}$ and $d\xi^a$ for vector and covector basis because they do not convey the actual physical meaning of these objects as they are used in physics.
+This is another instance of the math-for-mathematician issue.}
 
-So the situation is that we have a single clear physical object we want to study, but different and somewhat disconnected mathematical frameworks to study it.
-This should not be surprising as the frameworks are defined by mathematicians, as we said before, to solve their own needs.
-But it is an issue when these are used in physics without careful consideration as each mathematical framework only partially captures our physical understanding.
-That is why we argue that simply examining the mathematical structure used in a physical theory is not enough to gain a complete picture.
+So the situation is that we have a single clear physical object we want to study, a statistical distribution, but different and somewhat disconnected mathematical frameworks to study it.
+As we noted in the introduction, this should not be surprising as the frameworks are defined by mathematicians to solve their own needs.
+But it is an issue when these are used in physics without careful consideration as each mathematical framework may only partially capture our physical understanding.
+And this is precisely why we argue that simply examining the mathematical structure used in a physical theory is not enough to gain a complete picture: mathematics can only check for self-consistency and not for completeness or correctness.
+What we will need to do, in fact, is use elements from all these frameworks as some insights are better understood in one context and some in another.
 
@@
 
-Suppose we have a system which we can think of as made of infinitesimally small parts, which we call particles.
-Suppose  that each infinitesimal part can be assigned a state from a state space $\mathcal{S}$.
-The state of the whole system will be a distribution over those states.
+We now turn our attention to what we call infinitesimally reducible systems: a system whose overall state can be reduced to the state of its infinitesimal parts.
+That is, we can think of our system as made of infinitesimal pieces, which we call particles.
+Each particle is in a state from a state space $\mathcal{S}$.
+The state of the whole system, then, will be described by a distribution $\rho : \mathcal{S} \to \mathbb{R}$ over those states.
+That is, the elements of our distributions are the particles, the variable associate to each element is the state of the particle.
+Here we have the first problem: states, to be physically objective, must be coordinate independent objects.
+We \emph{do} want the density to be define on the actual space $\mathcal{S}$ or our distribution would not be physically meaningful.
+But we said before that, in general, densities are defined on the state variables, not on the object the state variables identify.
+
 Now suppose that this system evolves in a way that is deterministic and reversible, that is the state of a particle at one time is associated with one and only one state at another time.
 What does that tell us about our distribution?
-
 If the evolution is deterministic and reversible, then all the parts that are in one state will be mapped to another: the density should be exactly mapped to a different state and therefore should not change value.
 If the evolution is deterministic and reversible, then if I give you enough information to identify the state of an element at one time I also gave you enough information to identify the state of the same element at a future time: the information entropy will remain the same over time.
-And here lies the problem: how can we make sure that these quantities are preserved over time if they are not even observer independent?
+And here lies another problem: how can we make sure that these quantities are preserved over time if they are not even observer independent?
 How can we know if the density stays the same if we can't even compare the value of the densities at two different elements of the distribution?
@@
 
-This leads us to the following insight: \textbf{if we want to define deterministic and reversible evolution over distributions, they must be invariant distributions.
+We are forced to conclude that \textbf{if we want to define deterministic and reversible evolution over distributions, they must be invariant distributions.
 That is, the densities and information entropy must be independent of the choice of coordinates.} As we noted before, requiring one to be coordinate invariant will automatically require the other to be invariant as well.
+In other words, the state for an infinitesimally reducible system must be an invariant distribution over the states of the particles.
 
@@
 Our state space $\mathcal{S}$ must allow us to write densities whose value does not change under coordinate transformation.
+That is, the density must be change with the Jacobian (i.e.
+something we can integrate on) but also not change at all (i.e.
+the number is coordinate independent).
 This seems to be impossible given what we saw before: how can it work?
@@
 
-If we have a set of variables fixed by a single choice of unit then what we discussed before is the only possible way.
+Note how we have to separate the idea of state variables, a value used to identify a state, from a coordinate, a state variable that defines the system of units and references.
+The insight is that the link does not necessarily have to be one to one.
+One choice of coordinate can influence more then one state variables.
+But how many can it influence?
+
+It turns out that if we have a set of variables fixed by a single choice of unit, a single coordinate, then what we discussed before is the only possible way.
 Let $q$ be the coordinate, the variable that defines the unit.
-Suppose we have $n$ variable $\xi^a$ such that $\xi^1 = q$.
+Suppose we have $n$ variables $\xi^a$ such that $\xi^1 = q$.
 Each vector $v^a$ tangent to that space has $n$-components.
@@
 It is no wonder the link gets lost.
+But that's what, ultimately, is describing.
 
@@
 
-Let's set $\hbar$ as the unit we use to describe the amount of possible states (or the inverse of a unitary density). We have:
+We are now ready to see how conservation of entropy leads to Hamiltonian mechanics.
+For now, we assume our invariant distribution is characterized by only one coordinate which means two state variables $(q, k)$.
+We need to choose the units over which our density is defined.
+As a normalized density in space is defined over a volume (e.g.
+units of $1/m^3$), a normalized density for particle states will be defined over a region of possible states.
+By convention, let's set $\hbar$ as the unit we use to describe the amount of possible states in a region.
+That is, our   We have:
 \begin{equation}
-m_V = M \mu(V) =\int_V M \rho(q, k) \hbar dq dk = \int_V M \rho(q, p) dq dp
+m_V = M \mu(V) =\int_V \frac{m(q, k)}{\hbar} \hbar dq dk = \int_V M \rho(q, p) dq dp
 \end{equation}
@@
 
-We can think of the evolution as a vector field on $\mathcal{S}$ of components $S = (S_q, S_p) = (\frac{dq}{dt}, \frac{dp}{dt})$ that gives the direction in which states move in time.
+Note that we have our cake, the density transforms like a Jacobian, and we can eat it too, the Jacobian is one and therefore the density and the information entropy is invariant.
+This is precisely what we needed.
+But in so doing we have given a precise meaning to all the pieces: the wave number is the additional state variable that each coordinate must introduce to give invariant densities, $\hbar$ is the unit we use to count the possible states over a region which is the inverse of the unit of the invariant density, canonical variables are those that define the density in the proper units, a canonical transformation is one that leave the density invariant and each coordinate transformation is a canonical transformation because conjugate momentum must change like a covector.
+It is striking to us of much we were able to get with such a simple starting point.
+
+Now it's time to find the equations of motion.
+We can think of the evolution as a vector field on $\mathcal{S}$ of components $S = (S^q, S^p) = (\frac{dq}{dt}, \frac{dp}{dt})$ that gives the direction in which states move in time.
+After an infinitesimal time step, we'd have:
+\begin{equation}
+\begin{aligned}
+q(t+dt) &= q(t) + \frac{dq}{dt} dt = q(t) + S^q dt \\
+p(t+dt) &= p(t) + \frac{dp}{dt} dt = p(t) + S^p dt
+\end{aligned}
+\end{equation}
 The Jacobian for time evolution will be
@@
 |J| &= \left| \begin{matrix}
-1 + \dfrac{\partial S_q}{\partial q}dt & \dfrac{\partial S_q}{\partial p} dt \\[2.2ex]
-\dfrac{\partial S_p}{\partial q}  dt & 1 + \dfrac{\partial S_p}{\partial p} dt \end{matrix} \right| \\
-&= 1 + \left[ \dfrac{\partial S_q}{\partial q} + \dfrac{\partial S_p}{\partial p} \right]dt + O(dt^2)\\
+1 + \dfrac{\partial S^q}{\partial q}dt & \dfrac{\partial S^q}{\partial p} dt \\[2.2ex]
+\dfrac{\partial S^p}{\partial q}  dt & 1 + \dfrac{\partial S^p}{\partial p} dt \end{matrix} \right| \\
+&= 1 + \left[ \dfrac{\partial S^q}{\partial q} + \dfrac{\partial S^p}{\partial p} \right]dt + O(dt^2)\\
 &= 1 + div(S)dt + O(dt^2)\\
@@
 It's just the simple realization that if we require distributions to have a physically objective reality they must be coordinate independent.
+The premise is much simpler than the math employed.
+We can reason intuitively on the premise.
+We can think about what it means and when it is applicable.
+
+Note how we haven't talked about mass, acceleration or forces.
+All those Newtonian concepts didn't appear.
+And since we were able to get phase space and the equations of motion without them, it means they don't play any significant role.
+Note how all the concepts we used, states and parts of a system, are automatically coordinate invariant.
+Note how it's clear what type of time evolution we are not described: those that are non-deterministic or non-reversible.
+Those that would concentrate or spread our distributions.
+
+Note how each mathematical object found its natural place.
+Note how we had to cross the boundaries of the different mathematical frameworks to find meaning as the current mathematical frameworks we have do not lend themselves to talk about these issues.
+Note, though, that our reasoning is logically necessary and therefore, in principle, we should be able to create physically motivated mathematical framework that would capture these ideas.
+We just cannot expect mathematicians to magically do it for us.
 
@@
 
-\begin{thebibliography}{0}
+\bibliographystyle{alpha}
 
-	\bibitem{North} Jill North; The "Structure" of Physics: A Case Study, The Journal of Philosophy 106, no. 2 (2009), Pages 57-88, http://www.jstor.org/stable/20620153.
-
-	\bibitem{Curiel}Erik Curiel; Classical Mechanics Is Lagrangian; It Is Not Hamiltonian, The British Journal for the Philosophy of Science, Volume 65, Issue 2, 1 June 2014, Pages 269–321, https://doi.org/10.1093/bjps/axs034
-
-	\bibitem{Barrett1}Thomas William Barrett; On the Structure of Classical Mechanics, The British Journal for the Philosophy of Science, Volume 66, Issue 4, 1 December 2015, Pages 801–828, https://doi.org/10.1093/bjps/axu005
-
-	\bibitem{Barrett2}Thomas William Barrett; Equivalent and Inequivalent Formulations of Classical Mechanics, The British Journal for the Philosophy of Science, , axy017, https://doi.org/10.1093/bjps/axy017
-
-	\bibitem{carc1} G. Carcassi, C. A. Aidala: \emph{Assumptions of Physics} (in preparation). http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.
-
-	\bibitem{carc2} G.
-Carcassi, C.
-A.
-Aidala, D.
-J.
-Baker and L.
-Bieri; From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics, Journal of Physics Communications, 2, 4, 045026, 2018.
-
-\end{thebibliography}
+\bibliography{bibliography}{}
 

### 2018-11-21 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Addressing Christine and Mario second pass

@@
 \usepackage{tikz}
+\usepackage{forest}
+\usetikzlibrary{trees}
 \usetikzlibrary{calc}
@@
 Over the past decade there has been a renewed interest in the underpinning of classical mechanics.
-Different attempts and arguments have been put forward to understand what are the most important features of the different mathematical formulations, whether they are equivalent or whether one is more fundamental than the other~\cite{North,Curiel,Barrett1,Barrett2}.
+Different attempts and arguments have been put forward to understand what are the most important features of the various mathematical formulations, whether they are equivalent or whether one is more fundamental than the other~\cite{North,Curiel,Barrett1,Barrett2}.
 We believe, though, that there are two general problems with these attempts.
@@
 Mathematical equivalence does not imply physical equivalence, which is especially problematic for \cite{Barrett2} since it focuses on the former.
-In other words: we can have the same physics with different math and the same math with different physics.
+In other words, we can have the same physics with different math and the same math with different physics.
 The connection between math and physics is many-to-many because mathematics captures only the syntax of a particular description without its semantics, which at that point cannot be recovered.
@@
 a transformation that does not change the equation of motion) such that any Hamiltonian system, in the new coordinates, has the motion of a free particle for a finite amount of time.
-As \cite{North} notes, only frame independent relationships should be taken as fundamental.
-Does that mean that the Hamiltonian itself is not a fundamental relationship, since it is coordinate dependent and all Hamiltonians are the same anyway?
-It is also known\cite{AllSystemsAreHam} that for any first order system of ordinary differential equations we can find an appropriate symplectic structure such that the system can be written as a Hamiltonian system.
+As \cite{North} notes, only frame-independent relationships should be taken as fundamental.
+Does that mean that the Hamiltonian itself is not fundamental because it is not invariant under canonical transformations?
+Does it mean that all Hamiltonian systems are the same system, since there is always a set of canonical coordinates in which $H=p^2/2m$?
+It is also known\cite{AllSystemsAreHam} that for any first-order system of ordinary differential equations we can find an appropriate symplectic structure such that the system can be written as a Hamiltonian system.
 Does that mean that, in the end, all systems are physically equivalent, regardless of what they describe?
-The point is, we can't know whether our findings about mathematical structures are physically significant if we do not have a precise link between the mathematical symbols and their physical meaning.
+The point is, we cannot know whether our findings about mathematical structures are physically significant if we do not have a precise link between the mathematical symbols and their physical meaning.
 
 This leads us to the second problem we believe the previous works may have: when trying to attach meaning the thinking and intuition is still Newtonian.
-This is particularly evident in \cite{Curiel} where we start from physical ideas to get to Lagrangian dynamics.
+This is particularly evident in \cite{Curiel} where physical ideas provide the starting point to get to Lagrangian dynamics.
 The starting point is that a classical system is one that, roughly, obeys Newton's second law and therefore we have, for a free particle, $\dot{x} = v$ and $\dot{v} = 0$.
@@
 This, again, leads to confusion and we may miss a key element: in the transition from Newtonian to Lagrangian/Hamiltonian mechanics we expand the reference frames in which the equations are valid (from inertial frames to all frames) at the expense of restricting the class of systems we can describe (e.g.
-dissipative forces in general require a modification of the Euler-Lagrange equation using the Rayleigh dissipation function).
+dissipative forces in general require a modification of the Euler-Lagrange equation typically using the Rayleigh dissipation function).
 The point is, Newtonian concepts may prevent us from fully understanding Hamiltonian and Lagrangian mechanics.
@@
 It should be clear that we do not fault the authors.
-We fault us, the physicists.
+We fault ourselves, the physicists.
 Even just a century ago, there was a push in our field to find fundamental principles or laws that could serve as a starting point for the different theories.
@@
 Nowadays, unfortunately, many of the current fundamental theories, including Lagrangian, Hamiltonian, quantum mechanics and quantum field theory take as their starting point a particular mathematical structure.
-That is, we do not know what the value of a Lagrangian means or what the symplectic form represents: we just take them as given.
+That is, we do not know what the value of a Lagrangian means or what the symplectic form represents; we just take them as given.
 Moreover, a lot of work in theoretical physics nowadays starts by exploring novel abstract mathematical structures and then trying to see whether some physical meaning or prediction can be extracted.
@@
 
-We begin in section 2 by giving a brief summary about statistical distributions in general.
+We begin in section 2 by giving a brief summary of statistical distributions in general.
 These are your typical mathematical tools used to keep track of how charge is distributed in space or how a population of individuals is distributed in age.
@@
 information about the system at one time is equivalent to information about the system before or after).
-The requirement of coordinate independent distributions gives us the symplectic structure (i.e.
+The requirement of coordinate-independent distributions gives us the symplectic structure (i.e.
 conjugate pairs of position and momentum) while the requirement of conservation of information entropy gives us Hamiltonian evolution.
-In section 4 we discuss the result and provide a dictionary that allows us to give precise meaning to the different mathematical symbols, addressing the issues we highlighted before.
+In section 4 we discuss the result and provide a dictionary that allows us to give precise meaning to the different mathematical symbols, addressing the issues highlighted before.
 We also show how this understanding allows us to construct a classical analogue of the uncertainty principle.
@@
 \end{equation}
-The integral is used when some of the variables $\xi^a$ are continuous and a simple sum would be used over the discrete ones. We can also define the normalized density as:
+The integral is used when some of the variables $\xi^a$ are continuous, and a simple sum would be used over the discrete ones. We can also define the normalized density as:
 \begin{equation}
@@
 
-Before proceeding, though, let's give a quick introduction so that we have a clear understanding about what information entropy represents.
-In the physics community, for example, the concept is often vaguely referred to as ``knowledge" (and sometimes as ``lack of knowledge") which is incorrect and misleading.\footnote{We believe this characterization may be due to Jaines\cite{Jaines} who first introduced the concept of information theory within thermodynamics.} Information entropy was introduced by Shannon\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and as such should be treated.
-What he needed was a way to measure how much information a particular channel is able to transfer and therefore he needed a way to quantify precisely information as number of bits.
+Before proceeding, though, let us give a quick introduction so that we have a clear understanding about what information entropy represents.
+In the physics community, for example, the concept is often vaguely referred to as ``knowledge" (and sometimes as ``lack of knowledge") which is incorrect and misleading.\footnote{We believe this characterization may be due to Jaynes\cite{Jaynes} who first introduced the concept of information theory within thermodynamics.
+Note that we are only going to talk about information entropy and not thermodynamic entropy.
+We do believe there is a link between the two, but it is not the one Jaynes gives.
+We leave that discussion for another work.} Information entropy was introduced by Shannon\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and should be treated as such.
+What he needed was a way to measure how much information a particular channel is able to transfer and therefore a way to quantify precisely information as number of bits.
 So the idea is that we have a source that sends messages to a destination, and the message is encoded in zeros and ones.
 Information entropy is the optimal average number of zeros and ones that must be used to encode the messages.
-Below, the message is crippled and above the message is redundant.
+A message with fewer bits is crippled, and a message with more bits is redundant.
 
-To understand how this work in our context, let's go back to our distribution $\rho(\xi^a)$ and refine our thought experiment.
-Suppose we have a distribution, known to both us.
-Suppose you pick an element at random and I need to know the value of the variable for the element you picked.
-That is, you are the source, I am the destination and the message is the value of the element you picked from the distribution.
+To understand how this works in our context, let us go back to our distribution $\rho(\xi^a)$ and refine our thought experiment.
+Suppose two individuals, Alice and Bob, have full knowledge of the same distribution.
+Suppose Alice picks an element at random and Bob needs to know the value of the variable for the element Alice picked.
+That is, Alice is the source, Bob is the destination and the message is the value of the element Alice picked from the distribution.
 How long, on average, should the message be?
 How many ones and zeros?
-How many yes or no questions must I ask you before I know the value?
+How many yes or no questions must Bob ask Alice before he knows the value?
 For example, suppose there are 40 balls of which 20 are red and 20 are green.
@@
 Suppose there are 40 balls of which 10 red, 10 green, 10 blue and 10 yellow.
-In this case, two questions would suffice as two yes or no questions, two bits of information, can distinguish between 4 cases.
-In general, if $I$ is the number of questions, the number of bits, then $C = 2 ^ I$ are the number of cases you can distinguish with those questions.
+In this case, two questions would suffice as shown in figure \ref{twoQuestions}, since two yes or no questions, two bits of information, can distinguish 4 cases.
+
+\begin{figure}\label{twoQuestions}
+	\centering
+\begin{tikzpicture}[level distance=1.5cm,
+level 1/.style={sibling distance=6cm},
+level 2/.style={sibling distance=3cm}, sloped]
+\node {Is it red or green?}
+child {node {Is it red?}
+	child {
+		node {red - 11}
+		edge from parent node[pos=0.6, align=center, above] {Yes}
+	}
+	child {
+		node {green - 10}
+		edge from parent node[pos=0.6, align=center, above] {No}
+	}
+	edge from parent node[pos=0.6, align=center, above] {Yes}
+}
+child {node {Is it blue?}
+	child {
+		node {blue - 01}
+		edge from parent node[pos=0.6, align=center, above] {Yes}
+	}
+	child {
+		node {yellow - 11}
+		edge from parent node[pos=0.6, align=center, above] {No}
+	}
+	edge from parent node[pos=0.6, align=center, above] {No}
+};
+\end{tikzpicture}
+\caption{The decision tree for the example in the text.
+The nodes of the tree represent the questions Bob can pose to Alice and the bottom line shows the four different cases with the corresponding encodings.}
+\end{figure}
+
+In general, if $I$ is the number of questions, the number of bits, then $C = 2 ^ I$ are the number of cases one can distinguish with those questions.
 On the other hand, suppose for a particular value $\xi_0$ of a discrete variable we have $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
@@
 which is the information entropy of the distribution.
-That tells us the average number of bits you need to put into a message to tell me the value of the element of the distribution you picked at random.
+That tells us the average number of bits one needs to put into a message to tell someone else the value of the element of the distribution that was picked at random.
 
@@
 However, in the discussion we will refer to it as measured in bits, as we believe that the majority of the readers is more comfortable with the latter unit.
-Also, technically, that for continuous variables the expression has a slightly different meaning: the uniform distribution in a unit volume is assigned zero entropy so $I(\rho(\xi^a))$ represents the relative number of bits to achieve the same level of precision, which can be negative.
-That is, if we take a real number from a uniform distribution between 0 and $\frac{1}{2}$ we need to loose a bit of information to get to the same precision of a unitary distribution.
+Also, technically, for continuous variables the expression has a slightly different meaning: the uniform distribution in a unit volume is assigned zero entropy so $I(\rho(\xi^a))$ represents the relative number of bits to achieve the same level of precision, which can be negative.
+That is, if we take a real number from a uniform distribution between 0 and $\frac{1}{2}$ we need to lose a bit of information to get to the same precision of a unitary distribution.
 
@@
 What is also important is that it is a precisely defined quantity and that there is only one way to define it.
-It is the only expression that satisfies three simple requirements (continuity, monotonicity and additivity) that are necessary to measure information and that is how Shannon originally derived it.
-Insofar that we want to talk about amounts of information we have only one way to do it.
+It is the only expression that satisfies three simple requirements (continuity, monotonicity and additivity) that are necessary to measure information, and that is how Shannon originally derived it.
+Insofar that we want to talk about amounts of information we have only one way to do so.
 
@@
 Intuitively, if we change units the density changes value.
-For example, $1$ kg/m becomes $1000$ kg/km: the number changes and the units allow us to understand that the overall quantity remains the same but mathematics does not keep track of units.
+For example, a linear mass density of $1$ kg/m at $20$ Celsius for one references becomes a linear density of $1000$ kg/km at $20$ Celsius.
+The temperature does not change value (it is a scalar and therefore invariant) while the density changes value.
+Physically, we use the unit to understand that the overall quantity is the same.
+But mathematics only keeps track of the number, which changes, and is therefore coordinate dependent.
 In general, if $\hat{\xi}^a=\hat{\xi}^a(\xi^b)$ we can recover how the density changes by requiring that the total in each volume is the same no matter what variables are used.
@@
 
-To sum up, while we'd like to think of a density as a number associated to a point of a space, that's technically incorrect.
+To sum up, while we would like to think of a density as a number associated to a point of a space, that is technically incorrect.
 Given a manifold $\mathcal{M}$ upon which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \to \mathbb{R}$ in general.
-That would be a scalar function, invariant under change of variables, as a point is a point no matter how is it identified.
+That would be a scalar function, invariant under change of variables, as a point is a point no matter how it is identified.
 Instead we additionally need a set of functions $\xi^a : \mathcal{M} \to \mathbb{R}$ and only then we can define the density as $\rho : \mathbb{R}^n \to \mathbb{R}$.
-The density value is defined on the values for a set of variables, not at a point.\footnote{As the density is really a limit, we need to know how that limit is taken and the coordinates decide that.} That is, the density is undefined without a (differentiable) set of coordinates.
+The density value is defined on the values for a set of variables, not at a point.\footnote{As the density is really a limit, we need to know how that limit is taken, and the coordinates decide that.} That is, the density is undefined without a (differentiable) set of coordinates.
 
@@
 
-Different fields in math deal with this issue in different ways depending on their aims, which unfortunately leads to a fractured physical understanding given that one needs ideas and results from the different areas.
+Different fields in math deal with the coordinate dependency of densities in different ways depending on their aims, which unfortunately leads to a fractured physical understanding given that one needs ideas and results from the different areas.
 In statistics and information theory one simply accepts the transformation rules and does not try to create invariant objects.
@@
 In probability theory one starts with three objects: a set of outcomes $\mathcal{M}$, a set of events $\sigma_\mathcal{M}$ where each event is a collection of outcomes, and a probability measure $\mu : \sigma_\mathcal{M} \rightarrow [0,1]$ that assigns a number between zero and one to each event.
-That is, we don't assign a probability density to the points of $\mathcal{M}$ but we assign a finite probability to finite regions.
-We can then define a random variable $\xi : \mathcal{M} \rightarrow \mathbb{R}$ as a real valued function of the outcomes.
+That is, we do not assign a probability density to the points of $\mathcal{M}$ but we assign a finite probability to finite regions.
+We can then define a random variable $\xi : \mathcal{M} \rightarrow \mathbb{R}$ as a real-valued function of the outcomes.
 For each random variable we can define its cumulative distribution function $F_\xi(\xi_0)=\mu(\xi<\xi_0)$ and the probability density function is its Radon-–Nikodym derivative $\rho(\xi) = \frac{dF_\xi}{d\xi}$.
@@
 It also has the advantage that discrete and continuous quantities are treated in the same way.
-But in probability theory we have no notion of coordinate systems, vectors and all other geometrical objects.
+But in probability theory we have no notion of coordinate systems, vectors or any other geometrical objects.
 
@@
 
-This approach gives us a way to understand which objects are coordinate independent and which aren't.
+This approach gives us a way to understand which objects are coordinate independent and which are not.
 But it has two issues: it does not allow us to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.
@@
 So the situation is that we have a single clear physical object we want to study, a statistical distribution, but different and somewhat disconnected mathematical frameworks to study it.
-As we noted in the introduction, this should not be surprising as the frameworks are defined by mathematicians to solve their own needs.
-But it is an issue when these are used in physics without careful consideration as each mathematical framework may only partially capture our physical understanding.
+As noted in the introduction, this should not be surprising as the frameworks are defined by mathematicians to solve their own needs.
+But it is an issue when these are used in physics without careful consideration, as each mathematical framework may only partially capture our physical understanding.
 And this is precisely why we argue that simply examining the mathematical structure used in a physical theory is not enough to gain a complete picture: mathematics can only check for self-consistency and not for completeness or correctness.

### 2018-11-26 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Revised third section

@@
 
-As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work\cite{AoPPhy1} that aims to identify such starting points.
-Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
+As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work\cite{AoPPhy1} that aims to identify such starting points.\footnote{We should also mention \cite{Maudlin} as another attempt to formally derive mathematical structure from physical idea.} Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
 a symplectic structure where variables are organized in conjugate pairs).
@@
 Note that we are only going to talk about information entropy and not thermodynamic entropy.
-We do believe there is a link between the two, but it is not the one Jaynes gives.
-We leave that discussion for another work.} Information entropy was introduced by Shannon\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and should be treated as such.
+We do believe there is a link between the two, but it is not the one Jaynes gives which we believe to be unphysical.
+We leave that discussion for later works.} Information entropy was introduced by Shannon\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and should be treated as such.
 What he needed was a way to measure how much information a particular channel is able to transfer and therefore a way to quantify precisely information as number of bits.
@@
 	child {
-		node {yellow - 11}
+		node {yellow - 00}
 		edge from parent node[pos=0.6, align=center, above] {No}
@@
 \caption{The decision tree for the example in the text.
-The nodes of the tree represent the questions Bob can pose to Alice and the bottom line shows the four different cases with the corresponding encodings.}
+The nodes of the tree represent the questions Bob can pose to Alice to identify the color and the bottom line shows the four different cases with the corresponding encodings.}
 \end{figure}
@@
 Intuitively, if we change units the density changes value.
-For example, a linear mass density of $1$ kg/m at $20$ Celsius for one references becomes a linear density of $1000$ kg/km at $20$ Celsius.
+For example, a linear mass density of $1$ kg/m at $20$ Celsius for one reference becomes a linear density of $1000$ kg/km at $20$ Celsius.
 The temperature does not change value (it is a scalar and therefore invariant) while the density changes value.
@@
 
-\section{Coordinate independent distributions}
+\section{Coordinate-independent distributions}
 
-We now turn our attention to what we call infinitesimally reducible systems: a system whose overall state can be reduced to the state of its infinitesimal parts.
-That is, we can think of our system as made of infinitesimal pieces, which we call particles.
-Each particle is in a state from a state space $\mathcal{S}$.
-The state of the whole system, then, will be described by a distribution $\rho : \mathcal{S} \to \mathbb{R}$ over those states.
-That is, the elements of our distributions are the particles, the variable associate to each element is the state of the particle.
-Here we have the first problem: states, to be physically objective, must be coordinate independent objects.
-We \emph{do} want the density to be define on the actual space $\mathcal{S}$ or our distribution would not be physically meaningful.
-But we said before that, in general, densities are defined on the state variables, not on the object the state variables identify.
+We now turn our attention to what \cite{AoPPhy1} calls an infinitesimally reducible system.
+This is a system that can be thought as made of infinitesimally small parts, called particles, and for which giving a description of the full system is equivalent to giving a full description for all infinitesimal parts.
+Under this premise each particle is assigned a state from a state space $\mathcal{S}$ and the state of the whole system will be described by a distribution over those particle states.
+The elements of our distribution are the particles and the property that characterizes each element is its state.
 
-Now suppose that this system evolves in a way that is deterministic and reversible, that is the state of a particle at one time is associated with one and only one state at another time.
-What does that tell us about our distribution?
-If the evolution is deterministic and reversible, then all the parts that are in one state will be mapped to another: the density should be exactly mapped to a different state and therefore should not change value.
-If the evolution is deterministic and reversible, then if I give you enough information to identify the state of an element at one time I also gave you enough information to identify the state of the same element at a future time: the information entropy will remain the same over time.
-And here lies another problem: how can we make sure that these quantities are preserved over time if they are not even observer independent?
-How can we know if the density stays the same if we can't even compare the value of the densities at two different elements of the distribution?
-How can we tell whether the information entropy is conserved if we can't even define a unique value at a given time?
+Additionally, suppose that this system evolves in a way that is deterministic and reversible, that is, the state of a particle at one time is associated with one and only one state at another, future or past, time.\footnote{Note that by reversible we mean the ability to reconstruct the past, the dual of determinism, which we take it to mean the ability to predict the future.
+We do not mean a time symmetric process, where the equation of motion does not change when time is reversed, or a process that can be undone by another.} In this case, all the particles and only the particles that are in the same initial state will be mapped to the same final state: the density at the initial state must be the same as the density at the final state.
+Moreover, if Alice gave Bob enough information to identify the state of an element at one time, Alice also gave Bob enough information to identify the state of the same element at a future or past time: the information entropy will remain the same over time.
 
-We are forced to conclude that \textbf{if we want to define deterministic and reversible evolution over distributions, they must be invariant distributions.
-That is, the densities and information entropy must be independent of the choice of coordinates.} As we noted before, requiring one to be coordinate invariant will automatically require the other to be invariant as well.
-In other words, the state for an infinitesimally reducible system must be an invariant distribution over the states of the particles.
+While the physical setting is clear enough, it introduces a puzzle.
+States will be themselves identified by quantities, which we call state variables.
+But for states to be physically objective, they must be independent of reference frame.
+For example, if $A$ and $B$ are two reference frames and $\xi^a$ and $\hat{\xi}^b$ are the state variables one would use in the respective reference frames, we must have $s(\xi^a) = s(\hat{\xi}^b)$.
+Therefore, since the distribution is defined on the state itself, we must have $\rho(s(\xi^a))=\rho(s(\hat{\xi}^b))$.
+In this case, it seems we \emph{do} want the density to be defined on the actual space $\mathcal{S}$ and it seems it will have to be invariant under an arbitrary change of state variables.
+Accordingly, if we want densities and information entropy to be conserved over time, we first need to make sure these quantity are defined in a way that is observer independent, since otherwise they would be conserved in one frame but not in another.
+But this seems in apparent contradiction with the previous section and we need to clarify that.
 
-If our variables are discrete, all our integrals are sums and all our definitions work.
-Everything is already coordinate invariant.
-If our variables are continuous, instead, we have a problem.
-Because the densities and the information entropy vary with the Jacobian, we need something more.
-Our state space $\mathcal{S}$ must allow us to write densities whose value does not change under coordinate transformation.
-That is, the density must be change with the Jacobian (i.e.
+For starters, we concede that \textbf{only invariant distributions, those for which the density and information entropy do not change under a change of coordinates, can characterize the state of a system made of infinitesimal parts and its deterministic and reversible evolution.} Since requiring information entropy to be invariant will automatically require the density to be invariant as well and vice-versa, the requirement is one and the same.
+
+Now, if our variables are discrete, all our integrals are sums and all our definitions work: everything is already invariant.
+On the other hand, if our variables are continuous (i.e.
+the state space is a manifold) we have a problem as densities and information entropy will not be invariant under an arbitrary change of state variables.
+Our state space $\mathcal{S}$, somehow, must allow us to write densities that, under coordinate transformations, change with the Jacobian (i.e.
 something we can integrate on) but also not change at all (i.e.
-the number is coordinate independent).
-This seems to be impossible given what we saw before: how can it work?
+something that is coordinate independent).
+How can this possibly work?
 
@@
 We'll have $d\hat{q} d\hat{k} = dq dk$.
-Therefore we'll also have $\rho(q,k) = \rho(\hat{q}, \hat{k})$: the density is invariant and so is the information entropy.
+Therefore we will also have $\rho(q,k) = \rho(\hat{q}, \hat{k})$: the density is invariant and so is the information entropy.
 
-Note how we have to separate the idea of state variables, a value used to identify a state, from a coordinate, a state variable that defines the system of units and references.
-The insight is that the link does not necessarily have to be one to one.
-One choice of coordinate can influence more then one state variables.
+And here lies the solution to the puzzle: there is a difference between state variables, quantities used to identify a state, and coordinates, state variables that additionally define the system of reference and units.
+The density and information entropy must be invariant under arbitrary change of coordinates, not arbitrary change of state variables.
+The insight is that one choice of coordinate can influence more then one state variable.
 But how many can it influence?
 
-It turns out that if we have a set of variables fixed by a single choice of unit, a single coordinate, then what we discussed before is the only possible way.
+It turns out that if we have a set of variables fixed by a single choice of unit, a single coordinate, then a pair $(q, k)$ is the only possible way.
 Let $q$ be the coordinate, the variable that defines the unit.
 Suppose we have $n$ variables $\xi^a$ such that $\xi^1 = q$.
-Each vector $v^a$ tangent to that space has $n$-components.
-If we change unit $\hat{q}=\hat{q}(q)$ we have two constraints: the one unit change and the unitary Jacobian, and these must be enough to determine how the vector components change.
-The number of components must be equal to the number of constraints and therefore there are exactly two variables $\xi^a = (q, k)$.
+The vector $d\xi^a$ representing the infinitesimal displacement has $n$-components.
+If we change unit $\hat{q}=\hat{q}(q)$ we have two constraints: the unit change and the unitary Jacobian, and these must be enough to determine how the vector components change.
+If not, we would need to specify another relationship between the old $\xi^a$ and the new variables $\hat{\xi}^b$ which means it is not true that the units of all variables are fixed by a single choice of unit for $q$.
+As the linear system for $d\hat{\xi}^b$ must be solvable, the number of vector components must be equal to the number of constraints and therefore there are exactly two variables $\xi^a = (q, k)$.
 Then $k$ is the conjungate variable of $q$.
+There is no other choice.
 
-Physically, the coordinate plays a double role.
-It both defines a unit and it helps identify particle states.
-It is this double role that gives it an important character because a change in $q$ must induce a change in $k$ as their units and the unit of $\rho$ are not independent.
-Symplectic geometry strips away the units and does not elevate densities as primary objects.
-It is no wonder the link gets lost.
-But that's what, ultimately, is describing.
+To sum up, the density and information entropy will be invariant under coordinate transformations but will not be, in general, invariant under all state variable transformations.
+Physically, the coordinates play a double role.
+They define the reference frame and also help to identify particle states.
+It is this double role that gives them a special character because a change in $q$ must induce a change in $k$ as their units and the unit of $\rho$ are not independent.
+This is the key physical insight that will lead us to Hamiltonian mechanics.
+Note that symplectic geometry strips away the units and does not elevate densities as primary objects.
+It is no wonder this insight gets lost.
+Which goes back to our original point: we need to know what the mathematical structures represent physically and that cannot be deduced from the math itself.
 
@@
 
-We are now ready to see how conservation of entropy leads to Hamiltonian mechanics.
+We are now ready to see how conservation of entropy leads to the phase space of Hamiltonian mechanics.
 For now, we assume our invariant distribution is characterized by only one coordinate which means two state variables $(q, k)$.
-We need to choose the units over which our density is defined.
-As a normalized density in space is defined over a volume (e.g.
+While we have discussed units for the state variables, we still need to choose the unit for the density itself.
+In the same way a normalized density in space is a pure number over over a volume (e.g.
 units of $1/m^3$), a normalized density for particle states will be defined over a region of possible states.
-By convention, let's set $\hbar$ as the unit we use to describe the amount of possible states in a region.
-That is, our   We have:
+By convention, we use $\hbar$ as the unit we use to describe the amount of possible states in a region.
+We have:
 \begin{equation}
-m_V = M \mu(V) =\int_V \frac{m(q, k)}{\hbar} \hbar dq dk = \int_V M \rho(q, p) dq dp
+m_V = M \mu(V) =\int_V m(q, k) \hbar dq dk = \int_V M \rho(q, p) dq dp
 \end{equation}
-where we set $p=\hbar k$. We should recognize $p$ as conjugate momentum and $k$ the classical analogue of the wave number. Suppose we change variables. In general the Jacobian will be given by:
+where we set $p=\hbar k$. We recognize $p$ as conjugate momentum and $k$ the classical analogue of the wave number.
+
+Now suppose we change variables. In general the Jacobian will be given by:
 \begin{equation}
@@
 \end{equation}
-which we recognize to be the Poisson bracket. If the change of variable is simply a change of coordinates, a change of units, we have:
+which we recognize to be the Poisson bracket. If the change of variables is simply a change of coordinates, we have:
 \begin{equation}
@@
 \end{equation}
-Conjugate momentum transforms as a covariant component. Under coordinate transformations we have:
+We find that conjugate momentum transforms as a covariant component of a covector and that the density and information entropy are invariant:
 \begin{equation}
@@
 Note that we have our cake, the density transforms like a Jacobian, and we can eat it too, the Jacobian is one and therefore the density and the information entropy is invariant.
-This is precisely what we needed.
-But in so doing we have given a precise meaning to all the pieces: the wave number is the additional state variable that each coordinate must introduce to give invariant densities, $\hbar$ is the unit we use to count the possible states over a region which is the inverse of the unit of the invariant density, canonical variables are those that define the density in the proper units, a canonical transformation is one that leave the density invariant and each coordinate transformation is a canonical transformation because conjugate momentum must change like a covector.
-It is striking to us of much we were able to get with such a simple starting point.
+This is precisely what we needed and we have also given a precise meaning to all the pieces: the wave number $k$ is the additional state variable that each coordinate must introduce to give invariant densities, $\hbar$ is the unit we use to count the possible states over a region which is the inverse of the unit of the invariant density, canonical variables are those that define the density in the proper unit, a canonical transformation is one that leaves the density invariant and each coordinate transformation is a canonical transformation because conjugate momentum changes like a covector.
+It is striking how much we were able to get with such a simple starting point of coordinate-independent distributions.
 
-Now it's time to find the equations of motion.
+Now it is time to find the equations of motion.
 We can think of the evolution as a vector field on $\mathcal{S}$ of components $S = (S^q, S^p) = (\frac{dq}{dt}, \frac{dp}{dt})$ that gives the direction in which states move in time.
@@
 
-If we want the time evolution to transport the densities and to conserve information entropy, the Jacobian must be unitary and therefore $S$ is divergence-free and admits a potential. We have
+If we want the time evolution to transport  densities and to conserve information entropy, the Jacobian must be unitary and therefore $S$ is divergence-free and admits a potential. We have
 \begin{equation}
@@
 and we recognize Hamilton's equations.
-Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it's not just that the density and the areas are conserved in time: it is that they are the same for all canonical coordinates.
-That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is precisely the evolution of (classical) distributions that preserve information entropy and densities}.
+Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it is not just that the density and the areas are conserved in time: it is that they are the same for all canonical transformation including change of coordinates.
+That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is precisely the evolution of (classical) distributions over a manifold that preserve information entropy and densities}.
 
@@
 And the premise is not about cotangent bundles of Lie groups of quasi-q spaces on an infinite-dimensional lattice.
-It's just the simple realization that if we require distributions to have a physically objective reality they must be coordinate independent.
+It is just the simple realization that if we require distributions to have a physically objective reality they must be coordinate independent.
 The premise is much simpler than the math employed.
@@
 We can think about what it means and when it is applicable.
+We know what we are talking about.
 
-Note how we haven't talked about mass, acceleration or forces.
-All those Newtonian concepts didn't appear.
-And since we were able to get phase space and the equations of motion without them, it means they don't play any significant role.
+Note how we have not talked about mass, acceleration or forces.
+All those Newtonian concepts did not appear.
+And since we were able to get phase space and the equations of motion without them, it means they cannot play any significant role.
 Note how all the concepts we used, states and parts of a system, are automatically coordinate invariant.
-Note how it's clear what type of time evolution we are not described: those that are non-deterministic or non-reversible.
-Those that would concentrate or spread our distributions.
+Note how it is clear what type of time evolution we are \emph{not} describing: those that are non-deterministic or non-reversible.
+Those that would concentrate or spread our distributions, those that would change the amount of information to describe a particle state at a set level of precision are ruled out.
 
 Note how each mathematical object found its natural place.
-Note how we had to cross the boundaries of the different mathematical frameworks to find meaning as the current mathematical frameworks we have do not lend themselves to talk about these issues.
-Note, though, that our reasoning is logically necessary and therefore, in principle, we should be able to create physically motivated mathematical framework that would capture these ideas.
+Note how we had to cross the boundaries of the different mathematical frameworks to find meaning as the current mathematical frameworks we have do not lend themselves to talk about issues of units and coordinate-invariance of information entropy.
+Note, though, that our reasoning is logically necessary and therefore, in principle, we should be able to create a physically motivated mathematical framework that would capture these ideas.
 We just cannot expect mathematicians to magically do it for us.
@@
 
-To understand what happens with multiple degrees of freedom, let's first look at change of variables of marginal distributions.
-Suppose we have a distribution $\rho(\xi^a)$ that depends on many coordinates, on many units.
-If the density is coordinate invariant, then it will be so even if we change one coordinate, say $\xi^1$.
+To understand what happens with multiple degrees of freedom, let us first look at change of variables applied to marginal distributions.
+Suppose we have a distribution $\rho(\xi^a)$ that depends on many coordinates, each with its unit.
+If the density is coordinate invariant, then it will be so even if we change only one coordinate, say $\xi^1$.
 As we saw before, then we can find the conjugate variable $\xi^2$ that changes in the opposite way.
@@
 where $\xi^b$ are the remaining coordinates.
-If the variables $\xi^b$ are truly independent from $\xi^1$ and $\xi^2$, then we must be able to change units independently as well.
-Which means $\rho(\xi^b)$ must be coordinate invariant.
+If the variables $\xi^b$ are truly independent from $\xi^1$ and $\xi^2$, then a change of coordinates in the first will not affect the others and vice-versa.
+This means $\rho(\xi^b)$ must be coordinate invariant.
+In terms of information theory, invariance of the marginal densities is equivalent to being able to assign a coordinate invariant information entropy to each of the marginal distributions.
 
-We can imagine proceeding iteratively, one unit at a time.
-We'll find that our variables are paired $(q^i, p_i)$  with the first ones defining the units and the second ones defining the invariant areas.
+We can imagine proceeding iteratively, integrating away one pair at a time.
+We'll find that all our variables are paired $(q^i, p_i)$  with the first ones defining the units and the second ones defining the invariant areas.
 We call each pair an independent degree of freedom as it constitutes an independent choice of unit.
@@
 We are going to call it coordinate space as it includes all variables and only the variables that define the units and the reference frame.
-The state space for the particles, then, is a set of coordinates plus an equal number of conjugate variables that vary like covector components, therefore we recognize $\mathcal{S}=\mathsf{T}^*\mathcal{Q}$ as the cotangent bundle.\footnote{This does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number.
-The state space is just a manifold and the structure it has comes from how units transform, not because the variables represent different objects.
-Unfortunately mathematical structures strip out the units, so that knowledge is lost and then resurfaces in other structures.
-Changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects: it is relabeling the same states much like changing spatial coordinates relabels positions.
-It is not, in general, a canonical transformation as the density will not correspond to the invariant one, and it is not simply a change of coordinates.
-But it is still a valid change of state variables.}
+The state space for the particles, then, is a set of coordinates plus an equal number of conjugate variables that vary like covector components.
 
-In terms of information theory, this is equivalent to being able to assign a coordinate invariant information entropy to each of the marginal distributions.
-This means that relationships between the entropy of the total and marginal distributions are preserved.
+The idea, then, is that a distribution is truly coordinate-independent only when the density and information entropy for all the possible marginal distributions on any combinations of degrees of freedom are invariant, not just for the joint distribution.
+
+%Now, the fact that we can use a tangent bundle to track our particle states does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number.
+The state space is just a manifold, a set of points identified by $n$ continuous variable, and the structure it has comes from how units transform, not because the variables identify separate objects.
+Unfortunately mathematical structures strip out the units, the knowledge of the double role of coordinates is lost and then a different mathematical concept is used, one that has the same mathematical features of our particle state space but an altogether different meaning and construction.
+Changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects from a cotangent bundle to a tangent bundle: it is relabeling the same states much like changing spatial coordinates relabels positions.
+This change of variables, though, is not in general a canonical transformation which means the density will not correspond to the invariant one.
+But it is still a valid change of state variables and in fact, giving that velocity is gauge invariant, a very useful one.
+It allows to write Poisson brackets and find invariants of motion that are readily physically meaningful\cite{Holten}.
+
+%We can summ
+
+% In fact, if $(q^1, p_1)$ and $(q^2, p_2)$ are two degrees of freedom and $\rho_1(q^1, p_1)$ and $\rho_2(q^2, p_2)$ are the respective marginal distributions, we have now four ways of characterizing that they are independent.
+%These characterizations are equivalent because they repre
+
+%This means that relationships between the entropy of the total and marginal distributions are preserved.
 For example, saying that the marginal distributions $\rho_1(q^1, p_1)$ and $\rho_2(q^2, p_2)$ are independent is equivalent to saying that the entropy of the joint distribution is equal to the sum of the entropy of the marginals: $I(\rho(q^1,q^2, p_1, p_2)) = I(\rho_1(q^1, p_1)) +I(\rho_2(q^2, p_2))$.\footnote{This is also equivalent to saying that the joint distribution is the product of the marginals $\rho(q^1,q^2, p_1, p_2) = \rho_1(q^1, p_1)\rho_2(q^2, p_2)$.} That is, giving both values is equivalent to giving one and then the other.
 This means we can think of these relationships and quantities as real physical properties assigned to the degree of freedom and not just to the coordinates.
+
+%\begin{description}
+%	\item \emph{Coordinate independence}. The unit of $q^1$ can be chosen independently of $q^2$.
+%	\item \emph{State count}. If $dq^1 dp_1$ and $dq^2 dp_2$ represent how many possible states the count of states in each degree of freedom then the total count across is the product 
+%	\item \emph{Statistical independence}. The joint distribution is the product of the marginals: $\rho(q^1,q^2, p_1, p_2) = \rho_1(q^1, p_1)\rho_2(q^2, p_2)$.
+%	\item \emph{Information independence}.
+The information entropy of the joint distribution is the sum of the information entropy of the marginals: $I(\rho(q^1,q^2, p_1, p_2)) = I(\rho_1(q^1, p_1)) +I(\rho_2(q^2, p_2))$.
+Intuitively this means that giving the value of all state variables is equivalent to giving the values of both degree of freedom one after the other.
+%\end{description}
+%These characterizations are equivalent: you can't have one without the others.
 
@@
 is coordinate invariant.
-In fact, creating a marginal distribution means giving a family of surfaces parameterized by a set of variables, that do not intersect and whose union is the whole space and then calculating the integral as a function of the variables.
+In fact, creating a marginal distribution means giving a foliation (a parametrized family of surfaces that do not intersect and whose union is the whole space) and then calculating the integral as a function of the parameters.
 
@@
 It corresponds to a two dimensional vector product within each degree of freedom while it corresponds to a scalar product across degrees of freedom.
-As $\omega$ is preserved during coordinate transformations, the area within each degree of freedom is preserved meaning we are preserving the marginal distribution as well.
+As $\omega$ is preserved during coordinate transformations, the area within each degree of freedom is coordinate independent meaning areas of phase space are invariant and so are the marginal densities and information entropies.
 Across degrees of freedom, instead, we are preserving the angle meaning two perpendicular degrees of freedom remain perpendicular.
-But here perpedicular means independent: the density within the four dimensional volume is the product of the densities on the sides.
-As these are preserved under coordinate transformation, these now can correspond to actual physical entities.
+But here perpendicularity means independence, so whether two degrees of freedoms are independent is now invariant as well.
+As these concepts are coordinate invariant, they correspond to actual physical objects that all observers can agree on.
+They are part of the objective physical reality being described.
+
+Now, the fact that we can use a symplectic manifolds, or tangent bundles, to track our particle states does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number, defined on the coordinate space.
+The state space comes as one whole manifold, a set of points identified by $n$ continuous variable $\xi^a$, and the structure it has comes from how units transform, not because the variables identify separate objects.
+What really happen is that, unfortunately, mathematical structures strip out the units, the knowledge of the double role of coordinates is lost and then a different mathematical concept is used, one that has the same mathematical features of our particle state space but an altogether different meaning and construction.
+In the same vein, changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects from a cotangent bundle to a tangent bundle: it is simply relabeling the same states much like changing spatial coordinates relabels positions.
+This change of variables, though, is not in general a canonical transformation which means it does not preserve $\omega$ so the density will not correspond to the invariant one.
+But it is still a valid change of state variables and in fact, giving that velocity is gauge invariant, a very useful one.
+It allows to write Poisson brackets and find invariants of motion that are readily physically meaningful\cite{Holten}.
 
@@
 
-Deterministic and reversible evolution will preserve the densities and therefore preserve $\omega$.
+As deterministic and reversible evolution will conserve densities and information entropies, it will conserve $\omega$.
 That is, $\omega(v, w) = \omega(v',w')$ where $v'$ and $w'$ are the vectors evolved in time.
-In math terms, deterministic and reversible evolution for a distribution is a symplectomorphism.
+In math terms, deterministic and reversible evolution for an invariant distribution is a symplectomorphism.
 As before, $S = \left(\frac{d\xi^a}{dt}\right)$ is the vector field corresponding to the evolution.
-The vector components change according to $v'^a = \partial_b \xi^a(t+dt) v^b = (\delta^a_b + \partial_b S^a dt) v^b$.
+The vector components change according to the Jacobian.
+\begin{equation}
+	v'^a = \partial_b \xi^a(t+dt) v^b = (\delta^a_b + \partial_b S^a dt) v^b
+\end{equation}
 Therefore we have:
@@
 \partial_{a} S_{b} - \partial_{b} S_{a} &= curl(S_{a}) = 0 \\
-S_{a} &= \partial_{a}H \\
+S^{b} \omega_{ba} = S_{a} &= \partial_{a}H \\
 \frac{dq^i}{dt} &= \frac{\partial H}{\partial p_i}  \\
@@
 We recognize Hamilton's equations for multiple degrees of freedom.
-
-The symplectic structure was recovered by requiring invariant distributions which in turn requires invariant integration over independent degrees of freedom.
+Again, we stress how the symplectic structure was recovered by requiring invariant distributions which in turn requires invariant integration over independent degrees of freedom.
 Note how the requirement was given outside of the symplectic structure itself and therefore it cannot be recovered simply from that.
@@
 In other words, particles are not point-like in classical mechanics.
-They are infinitesimal cells of phase-space and that's why their evolution is symplectomorphism.
+They are infinitesimal cells of phase-space and that is why their evolution is symplectomorphism.
 
@@
 If one is convinced that that is not realistic then he is already convinced that classical mechanics will not hold in those regimes.
-And if that approximation fails, it's not that we need an altogether new set of rules: the macroscopic object is still the same.
+And if that approximation fails, it is not that we need an altogether new set of rules: the macroscopic object is still the same.
 We simply need to change how the subdivision works.
@@
 Among all possible choices of state variables, we conveniently choose the ones in which the distribution can be expressed as a density.
-That's the differentiable structure on the manifold.
+That is the differentiable structure on the manifold.
 The tangent and cotangent space are the tools to keep track of the process of infinitesimal subdivision and they are linear spaces because the space of all possible distributions is itself a linear space.
@@
 
-The larger point is that physics is one and it can't be understood piecemeal.
+The larger point is that physics is one and it cannot be understood piecemeal.
 Nature does not separate between dynamical systems and thermodynamics, between topologies and $\sigma$-algebras, between classical systems and quantum systems.
@@
 \end{align*}
-Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and its elements, and therefore it can't shrink the distribution more than a set amount.
+Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and its elements, and therefore it cannot shrink the distribution more than a set amount.
 If it did, it would start concentrating the density, meaning that multiple states have part of their density mapped to a single state and the dynamics would not be reversible.

### 2018-11-27 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Integrated Christine's comments on 3rd section

@@
 
-As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work\cite{AoPPhy1} that aims to identify such starting points.\footnote{We should also mention \cite{Maudlin} as another attempt to formally derive mathematical structure from physical idea.} Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
+As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work\cite{AoPPhy1} that aims to identify such starting points.\footnote{We should also mention \cite{Maudlin} as another attempt to formally derive mathematical structure from physical ideas.} Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
 a symplectic structure where variables are organized in conjugate pairs).
@@
 Note that we are only going to talk about information entropy and not thermodynamic entropy.
-We do believe there is a link between the two, but it is not the one Jaynes gives which we believe to be unphysical.
+We do believe there is a link between the two, but it is not the one Jaynes gives, which we believe to be unphysical.
 We leave that discussion for later works.} Information entropy was introduced by Shannon\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and should be treated as such.
@@
 Information entropy is the optimal average number of zeros and ones that must be used to encode the messages.
-A message with fewer bits is crippled, and a message with more bits is redundant.
+If the source sends fewer bits per message than required by the information entropy, some information will be lost.
+If the source sends more bits per message than required by the information entropy, some information will be redundant.
 
@@
 There is one feature about densities we need to be fully aware of: in general they are not invariant under change of variables.
-Intuitively, if we change units the density changes value.
-For example, a linear mass density of $1$ kg/m at $20$ Celsius for one reference becomes a linear density of $1000$ kg/km at $20$ Celsius.
-The temperature does not change value (it is a scalar and therefore invariant) while the density changes value.
+Intuitively, if we change units the density changes its numerical value.
+For example, $1$ kg/m becomes $1000$ kg/km.
 Physically, we use the unit to understand that the overall quantity is the same.
@@
 
-Additionally, suppose that this system evolves in a way that is deterministic and reversible, that is, the state of a particle at one time is associated with one and only one state at another, future or past, time.\footnote{Note that by reversible we mean the ability to reconstruct the past, the dual of determinism, which we take it to mean the ability to predict the future.
+Additionally, suppose that this system evolves in a way that is deterministic and reversible, that is, the state of a particle at one time is associated with one and only one state at another, future or past, time.\footnote{Note that by reversible we mean the ability to reconstruct the past, the dual of determinism, which we take to mean the ability to predict the future.
 We do not mean a time symmetric process, where the equation of motion does not change when time is reversed, or a process that can be undone by another.} In this case, all the particles and only the particles that are in the same initial state will be mapped to the same final state: the density at the initial state must be the same as the density at the final state.
@@
 In this case, it seems we \emph{do} want the density to be defined on the actual space $\mathcal{S}$ and it seems it will have to be invariant under an arbitrary change of state variables.
-Accordingly, if we want densities and information entropy to be conserved over time, we first need to make sure these quantity are defined in a way that is observer independent, since otherwise they would be conserved in one frame but not in another.
+Accordingly, if we want densities and information entropy to be conserved over time, we first need to make sure these quantities are defined in a way that is observer independent, since otherwise they would be conserved in one frame but not in another.
 But this seems in apparent contradiction with the previous section and we need to clarify that.
@@
 For example, $\hat{q}$ will be kilometers and $\hat{k}$ inverse kilometers.
-We'll have $d\hat{q} d\hat{k} = dq dk$.
+We will have $d\hat{q} d\hat{k} = dq dk$.
 Therefore we will also have $\rho(q,k) = \rho(\hat{q}, \hat{k})$: the density is invariant and so is the information entropy.
@@
 The vector $d\xi^a$ representing the infinitesimal displacement has $n$-components.
-If we change unit $\hat{q}=\hat{q}(q)$ we have two constraints: the unit change and the unitary Jacobian, and these must be enough to determine how the vector components change.
+If we change unit $\hat{q}=\hat{q}(q)$ we have two constraints: the unit change and density invariance, which means the Jacobian must be unitary.
+These must be enough to determine how the vector components change.
 If not, we would need to specify another relationship between the old $\xi^a$ and the new variables $\hat{\xi}^b$ which means it is not true that the units of all variables are fixed by a single choice of unit for $q$.
@@
 
-We are now ready to see how conservation of entropy leads to the phase space of Hamiltonian mechanics.
+We are now ready to see how conservation of information entropy leads to the phase space of Hamiltonian mechanics.
 For now, we assume our invariant distribution is characterized by only one coordinate which means two state variables $(q, k)$.
 While we have discussed units for the state variables, we still need to choose the unit for the density itself.
-In the same way a normalized density in space is a pure number over over a volume (e.g.
+In the same way a normalized density in space is a pure number over a volume (e.g.
 units of $1/m^3$), a normalized density for particle states will be defined over a region of possible states.
-By convention, we use $\hbar$ as the unit we use to describe the amount of possible states in a region.
+By convention, we use $\hbar$ as the unit to describe the amount of possible states in a region.
 We have:
@@
 
-Note that we have our cake, the density transforms like a Jacobian, and we can eat it too, the Jacobian is one and therefore the density and the information entropy is invariant.
-This is precisely what we needed and we have also given a precise meaning to all the pieces: the wave number $k$ is the additional state variable that each coordinate must introduce to give invariant densities, $\hbar$ is the unit we use to count the possible states over a region which is the inverse of the unit of the invariant density, canonical variables are those that define the density in the proper unit, a canonical transformation is one that leaves the density invariant and each coordinate transformation is a canonical transformation because conjugate momentum changes like a covector.
+Note that we have our cake, the density transforms like a Jacobian, and we can eat it too, the Jacobian is one and therefore the density and the information entropy are invariant.
+This is exactly what we needed and we have also given a precise meaning to all the pieces: the wave number $k$ is the additional state variable that each coordinate must introduce to give invariant densities, $\hbar$ is the unit we use to count the possible states over a region which is the inverse of the unit of the invariant density, canonical variables are those that define the density in the proper unit, a canonical transformation is one that leaves the density invariant and each coordinate transformation is a canonical transformation because conjugate momentum changes like a covector.
 It is striking how much we were able to get with such a simple starting point of coordinate-independent distributions.
@@
 We can think of the evolution as a vector field on $\mathcal{S}$ of components $S = (S^q, S^p) = (\frac{dq}{dt}, \frac{dp}{dt})$ that gives the direction in which states move in time.
-After an infinitesimal time step, we'd have:
+After an infinitesimal time step, we would have:
 \begin{equation}
@@
 and we recognize Hamilton's equations.
-Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it is not just that the density and the areas are conserved in time: it is that they are the same for all canonical transformation including change of coordinates.
-That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is precisely the evolution of (classical) distributions over a manifold that preserve information entropy and densities}.
+Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it is not just that the density and the areas are conserved in time: it is that they are the same for all canonical transformations including change of coordinates.
+That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is precisely the evolution of (classical) distributions over a manifold that preserves information entropy and densities}.
 
-We'd like to stress how, once the simple premise of invariant distributions is accepted, all elements of Hamiltonian mechanics are necessary and have a straightforward physical meaning.
+We would like to stress how, once the simple premise of invariant distributions under deterministic and reversible evolution is accepted, all elements of Hamiltonian mechanics are necessary and have a straightforward physical meaning.
 Not only that: because of the equivalence we know that everything in Hamiltonian mechanics can be understood with just that simple premise.
@@
 Note how we have not talked about mass, acceleration or forces.
-All those Newtonian concepts did not appear.
+None of those Newtonian concepts appeared.
 And since we were able to get phase space and the equations of motion without them, it means they cannot play any significant role.
@@
 We can imagine proceeding iteratively, integrating away one pair at a time.
-We'll find that all our variables are paired $(q^i, p_i)$  with the first ones defining the units and the second ones defining the invariant areas.
+We will find that all our variables are paired $(q^i, p_i)$  with the first ones defining the units and the second ones defining the invariant areas.
 We call each pair an independent degree of freedom as it constitutes an independent choice of unit.
@@
 
-The idea, then, is that a distribution is truly coordinate-independent only when the density and information entropy for all the possible marginal distributions on any combinations of degrees of freedom are invariant, not just for the joint distribution.
+The idea, then, is that a distribution is truly coordinate-independent only when the density and information entropy for all the possible marginal distributions on any combination of degrees of freedom are invariant, not just for the joint distribution.
 
@@
 To understand why, note that it corresponds to the case where $\rho(\xi^a)=1$.
-If the surface is a parallelopiped in some variables, each face that gives a positive contribution will have a parallel face that would give an equal and opposite contribution.
-As the closed integral needs to be zero for all parallelopipeds, it will also be zero for an arbitrary closed surface.
+If the surface is a parallelepiped in some variables, each face that gives a positive contribution will have a parallel face that would give an equal and opposite contribution.
+As the closed integral needs to be zero for all parallelepipeds, it will also be zero for an arbitrary closed surface.
 
 Under a change of variables, the two-form $\omega$ must be invariant.
-That is, phase space is a symplectic manifold.
-And, as any symplectic manifald, we can find Darboux variables $q^i, p_j$ and express the two-form as
+That is, the particle state space $(\mathcal{S}, \omega)$ is a symplectic manifold.
+And, as any symplectic manifold, we can find Darboux variables $q^i, p_j$ and express the two-form as
 \begin{equation}
@@
 Across degrees of freedom, instead, we are preserving the angle meaning two perpendicular degrees of freedom remain perpendicular.
-But here perpendicularity means independence, so whether two degrees of freedoms are independent is now invariant as well.
+But here perpendicularity means independence, so whether two degrees of freedom are independent is now invariant as well.
 As these concepts are coordinate invariant, they correspond to actual physical objects that all observers can agree on.
@@
 
-Now, the fact that we can use a symplectic manifolds, or tangent bundles, to track our particle states does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number, defined on the coordinate space.
-The state space comes as one whole manifold, a set of points identified by $n$ continuous variable $\xi^a$, and the structure it has comes from how units transform, not because the variables identify separate objects.
-What really happen is that, unfortunately, mathematical structures strip out the units, the knowledge of the double role of coordinates is lost and then a different mathematical concept is used, one that has the same mathematical features of our particle state space but an altogether different meaning and construction.
+Now, the fact that we can use symplectic manifolds, or tangent bundles, to track our particle states does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number, defined on the coordinate space.
+The state space comes as one whole manifold, a set of points identified by $n$ continuous variables $\xi^a$, and the structure it has comes from how units transform, not because the variables identify separate objects.
+Unfortunately, mathematical structures strip out the units, the knowledge of the double role of coordinates is lost and then a different mathematical concept is used, one that has the same mathematical features of our particle state space but an altogether different meaning and construction.
 In the same vein, changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects from a cotangent bundle to a tangent bundle: it is simply relabeling the same states much like changing spatial coordinates relabels positions.
 This change of variables, though, is not in general a canonical transformation which means it does not preserve $\omega$ so the density will not correspond to the invariant one.
-But it is still a valid change of state variables and in fact, giving that velocity is gauge invariant, a very useful one.
-It allows to write Poisson brackets and find invariants of motion that are readily physically meaningful\cite{Holten}.
+But it is still a valid change of state variables and in fact, given that velocity is gauge invariant, a very useful one.
+It allows one to write Poisson brackets and find invariants of motion that are readily physically meaningful\cite{Holten}.
 

### 2018-12-07 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Reviewed discussions and conclusions

@@
 \textbf{Hamiltonian mechanics is a type of statistical mechanics}, one with no probability and deterministic and reversible laws.
-The fact that entropy can be defined objectively and is conserved through evolution cannot be simply seen as a curiosity.
 
-To us, it does not make sense to start our physical discourse by imposing a state space for point-like particles, that somehow does not require only points but a particular differentiable structure as well.
-Then, if we are so keen, put distributions on it which turn out, somewhat magically, to have density and entropy defined in a coordinate independent way.
-It makes a lot more sense to start by saying we have a finite system which we pretend to be made of infinitesimal parts, which we keep dividing as we please and derive the state space of particles as the limit of such process of subdivision.
-This simultaneously gives the actual mathematical structure and its physical meaning.
+The usual approach posits point-like particles, with a particular state space and equations of motion.
+Then, coincidentally, finds this setting convenient for statistical mechanics because it has nice properties.
+In retrospect, this should sound backwards.
+We believe the approach presented here is preferable, first of all, because it justifies with a single insight the state space, the evolution and why that state space is so compatible with statistical mechanics.
 
-Not only is the second approach preferable because with fewer starting points it justifies the state space, the evolution and why that state space is so compatible with statistical mechanics, it also maps better to what we do in classical physics.
-We study composite systems (planets, cannonballs, beads on a wire, fluids) which we sometime pretend we can recursively divide.
+Second, it maps better to what we do in classical physics.
+We study composite systems (planets, cannonballs, beads on a wire, fluids) which we sometimes pretend we can recursively divide.
 None of those objects, in fact, has a single position or momentum: only the center of mass has a single position and momentum but those are precisely the expectation values $\int q \rho(q,p) dq dp$ and $\int p \rho(q,p) dq dp$ over the distribution.
-In other words, particles are not point-like in classical mechanics.
-They are infinitesimal cells of phase-space and that is why their evolution is symplectomorphism.
-
 Under this light, asking whether classical mechanics is tenable coincides with asking whether we can really keep dividing our object indefinitely.
-If one is convinced that that is not realistic then he is already convinced that classical mechanics will not hold in those regimes.
-And if that approximation fails, it is not that we need an altogether new set of rules: the macroscopic object is still the same.
+If one is convinced that this is not realistic then one is already convinced that classical mechanics will not hold in those regimes.
+But if that approximation fails, we do not need an altogether new set of ideas: the macroscopic object is still the same.
 We simply need to change how the subdivision works.
 
-Also note that this view tells us exactly what the mathematical framework of symplectic geometry is supposed to represent.
-The state space of particles is a manifold because we use real numbers to identify the states.
-Among all possible choices of state variables, we conveniently choose the ones in which the distribution can be expressed as a density.
-That is the differentiable structure on the manifold.
-The tangent and cotangent space are the tools to keep track of the process of infinitesimal subdivision and they are linear spaces because the space of all possible distributions is itself a linear space.
-Moreover the differentiable manifold has a symplectic structure because the density is invariant under coordinate transformation.
-Therefore among all possible choices of differentiable state variables we choose the canonical ones, the ones in which the density is expressed in appropriate units.
-Every requirement is physically motivated.
-No extra requirements are left.
+Third, this new approach is conceptually more compatible with the other physical theories.
+We have already seen how elements of special relativity (e.g.
+coordinate invariance, momentum as covector), statistical mechanics (e.g.
+joint and marginal distributions), quantum mechanics (e.g.
+wave number) and classical Hamiltonian mechanics (e.g.
+phase space, Poisson brackets) already emerged naturally.
+Furthermore, note that in general relativity point-like particles would be singularities that puncture space-time and would not move along a geodesics, while particles as infinitesimal parts would.
+Note that if the evolution is deterministic and reversible, it means the evolution of the system depends only on itself, which means that the system is isolated.
+That is: we have an association between isolated system and conservation of energy.\footnote{To be precise, conservation of energy is not an invariant property.
+The fact that a Hamiltonian exists, though, is an invariant property.} This sounds strikingly close to the first law of thermodynamics.
 
-While the power of this shift can really be appreciated in the context of our broader work, it should be clear how with relative few starting points, elements of relativity (e.g.
-coordinate invariance), statistical mechanics (e.g.
-physical distributions), quantum mechanics (e.g.
-wave number, conjugate quantities) and classical Hamiltonian mechanics (e.g.
-phase space, Poisson brackets) already emerge and form a coherent picture.
-In particular note how energy (in the form of the Hamiltonian) and entropy (in the form of information entropy) both come out from the single unassuming idea of invariant distributions.
-And note how conservation of entropy is related to deterministic and reversible evolution.
-Also note how deterministic and reversible evolution means that the system is only influenced by itself, which means it is not influenced by anything else, which means it is isolated, which is linked to energy conservation.\footnote{To be precise, conservation of energy is not an invariant property.
-The fact that a Hamiltonian exists, though, is an invariant property.} Surely this cannot be all coincidence.
-
-The larger point is that physics is one and it cannot be understood piecemeal.
+The larger point is that physics is one and it cannot truly be understood piecemeal.
 Nature does not separate between dynamical systems and thermodynamics, between topologies and $\sigma$-algebras, between classical systems and quantum systems.
@@
 
-Having a coherent picture across the different areas is not only intellectually more satisfying, but it allows us to switch perspective when the intuition from one fails.
-As a practical example, we can show how understanding the link between Hamiltonian mechanics and information theory leads very straightforwardly to a classical equivalent for the uncertainty principle.
+Having a coherent set of ideas, a coherent picture across the different areas is not only intellectually more satisfying, but it allows us to switch perspective when the intuition from one fails.
+As a practical example, we can show how understanding the link between Hamiltonian mechanics and information theory leads to a classical analogue for the uncertainty principle.
 
@@
 
-As we saw, Hamiltonian dynamics not only conserves the energy of all particles, but it also conserves the information entropy of the distribution as a whole.
-We can imagine that there is going to be some link between the spread of a distribution and the number of bits required to identify an element.
+As we saw, Hamiltonian dynamics not only conserves the energy of individual particles, but it also conserves the information entropy of the distribution as a whole.
+Intuitively, we can see that there is some link between the spread of a distribution and the number of bits required to identify an element.
+For example, if we have two uniform distributions with different ranges, then Alice will need to give Bob more information to identify the value from the bigger range.
 So we ask: what is the distribution that minimizes the spread given a certain amount of information entropy?
@@
 Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and its elements, and therefore it cannot shrink the distribution more than a set amount.
-If it did, it would start concentrating the density, meaning that multiple states have part of their density mapped to a single state and the dynamics would not be reversible.
+If it did, it would start concentrating the density, meaning that multiple states would be mapped to a single state and the dynamics would not be reversible.
+In this form $I_0$ is not a fundamental constant but a parameter that depends upon the initial condition.
+While this is already striking, \cite{AoPPhy1} shows a simple argument to turn this into something more.
 
-\section{Conclusion}
+Suppose that we have a distribution where the particles move non-derministically.
+That is, their motion is partly caused by some other agent in a way that can be characterized by random motion.
+Now suppose our distribution is at an equilibrium with that random motion.
+Then the particles change state randomly but the overall distribution remains the same.
+That is, the random motion simply reshuffles the elements.
+In this case the minimum achievable for the information entropy of the distribution depends on the random motion itself.
+The stronger it is, the more spread the distribution has to be, the higher the minimum threshold for the information entropy.
 
-In this paper we have shown how classical Hamiltonian mechanics coincides with conservation of information entropy.
-The structure of phase space is required to be able to define densities and entropy in a way that is coordinate invariant, and therefore physically meaningful.
-This clarifies the link between Hamiltonian mechanics, statistical mechanics and deterministic and reversible motion.
+Now suppose that source of chaotic motion is something intrinsic to a physical system, something it cannot be avoided.
+Maybe it is caused by interaction with the environment, from which the system can never be fully isolated.
+Or maybe it is caused by some internal motion, which can never be fully accessible.
+Or maybe it is caused by some background process with space itself, since the vacuum is not just empty space.
+The point here is not to identify the source of this non-deterministic motion: the point is to show that assuming one exists does not sound unreasonable.
 
-In the narrower context, it allows us to understand what physical object is represented by each mathematical one. This can be summarized in table \ref{dictionary}.
-\begin{table}[h]
+If one buys into the argument, one can posit that there exists an $I_\emptyset$ associated with this chaotic process that acts as the theoretical limit for the information entropy of our distributions.
+The farther the information entropy of our distribution is from that value, the more we can assume that the parts of the distribution evolve deterministically.
+The closer it is, the more our assumptions breaks down.
+This is what \cite{AoPPhy1} refers to as the classical uncertainty principle.
+
+\subsection*{A dictionary}
+
+We mentioned several time how this approach allows us to give precise physical meaning to the different mathematical objects.
+We leave in table \ref{dictionary} a summary of these connections in form of a math-to-physics dictionary.
+\begin{table}[h!]
 	\centering
@@
 \end{table}
-It also allows us to see why deterministic and reversible evolution must conserve entropy and why isolated systems must conserve energy.
+
+We can also better understand what physical requirement each mathematical structure is used for.
+Manifolds are used to represent a set of physical objects that can be identified by $n$ continuous variables.
+A differentiable structure on a manifold is used to define densities.
+Non-differentiable transformations, in fact, do not define a Jacobian and therefore cannot express the density in the new variables.
+A symplectic manifold is used to define coordinate invariant densities.
+Non-canonical transformations, in fact, have non unitary Jacobians and therefore the density would change the value of the density.
+
+\section{Conclusion}
+
+In this paper we have shown how classical Hamiltonian mechanics coincides with conservation of information entropy.
+The structure of phase space is required to be able to define densities and information entropy in a way that is coordinate invariant, and therefore physically meaningful.
+The equations of motion and the existence of the Hamiltonian are required to conserve densities and information entropy over time.
+This clarifies the link between Hamiltonian mechanics, statistical mechanics and deterministic and reversible motion.
+
+In the narrower context, it allows us to understand what physical object is represented by each mathematical one.
 In the broader context, it allows us to make deeper connections among different areas of math and physics, and pushes us to rethink our starting points.
@@
 
-These results represent one set of insights from our larger project, Assumptions of Physics, that aims to rederive the standard laws from a handful of physical starting points.
-We believe that a reorganization of fundamental physics on more physically meaningful conceptual footing and more rigorous mathematical grounds is not only possible, but long overdue.
+%These results represent one set of insights from our larger project, Assumptions of Physics, that aims to rederive the standard laws from a handful of physical starting points.
+
+This approach may point to a reorganization of fundamental physics on a more physically rigorous conceptual footing and more meaningful mathematical grounds which we believe is not only possible, but long overdue.
 

### 2018-12-12 Articles/TopologyAndExperimentalDistinguishability/article-topologyproceedings.tex
Minor changes after acceptance

@@
 \begin{rem}
-	Experimental observations are not closed under negation. The existence of an experimental test to verify a statement does not imply the existence of an experimental test to verify its negation.
+	Experimental observations are not closed under negation.
+The existence of an experimental test to verify a statement does not imply the existence of an experimental test to verify its negation.\footnote{Continuing the example, finding a black swan constitutes proof that black swans exist while not finding any black swans does not constitute proof that black swans do not exist.}
 \end{rem}
@@
 As experimental relationships are themselves distinguishable, we can recursively form experimental relationships between experimental relationships leading to functions of arbitrary order while remaining within the definitions provided.
-The framework is therefore complete.
+The framework is therefore closed.
 
@@
 In light of this work, the effectiveness of mathematics in the natural sciences is perhaps not so unreasonable.
-We hope that the methods and results shown here can provide a more solid foundation to formalize experimental sciences.
+We hope that the methods and results shown here can provide a more solid foundation to formalize experimental sciences.\footnote{For the latest developments in our effort to provide a more solid mathematical foundations to experimental sciences see \cite{Book}}
 
@@
 	\bibitem{russell} B. Russell, {\it The problems of philosophy}, Home University Library, 1912.
+
+	\smallskip
+
+	\bibitem{Book} G. Carcassi, C. A. Aidala: Assumptions of Physics (in preparation). http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.
 \end{thebibliography}

### 2018-12-12 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Final draft sent to everybody

@@
 	In this work we show the equivalence between Hamiltonian mechanics and conservation of information entropy.
-We will show that distributions with coordinate independent values for information entropy require that the manifold on which the distribution is defined is charted by conjugate pairs (i.e.
-it is a symplectic manifold).
+We will show that distributions with coordinate independent values for information entropy require that the manifold on which the distribution is defined is charted by conjugate pairs (i.e.~it is a symplectic manifold).
 We will also show that further requiring that the information entropy is conserved during the evolution yields Hamilton's equations.
@@
 This can be either a discrete quantity, like the number of people we are considering, or a continuous quantity, like the total amount of mass.
-We also have a set of $n$ variables $\xi^a$ over which the elements are distributed and a function $m(\xi^a)$ that gives us the amount for each value possible value such that:
+We also have a set of $n$ variables $\xi^a$ over which the elements are distributed and a function $m(\xi^a)$ that gives us the amount for each possible value such that:
 \begin{equation}
@@
 The normalized distribution $\rho(\xi^a)$, as defined, is a statistical distribution that represents the entirety of the system being described.
-Yet, we can give it an additional meaning of probability if we run the following thought experiment: suppose we took an element at random, each element with equal chance, what is the probability $P(\xi^a \in V)$ that the values for $\xi^a$ associated with that element are within a particular region $V$?
+Yet, we can give it an additional meaning of probability if we run the following thought experiment: suppose we took an element at random, each element with equal chance.
+That is, we have a uniform distribution on the elements (not the values).
+What is the probability $P(\xi^a \in V)$ that the values for $\xi^a$ associated with that element are within a particular region $V$?
 The answer must be $\mu(V)$ and therefore $\rho(\xi^a)$ is the probability density.
@@
 Therefore the information needed to identify a particular case is $I(\xi^a)=\log \frac{1}{\rho(\xi^a)}$.
-If we take the average and change the base of the logarithm, we have:
+If we take the expectation value of this expression over the distribution and change the base of the logarithm, we have:
 \begin{equation}
@@
 which is the information entropy of the distribution.
-That tells us the average number of bits one needs to put into a message to tell someone else the value of the element of the distribution that was picked at random.
+This represents the average number of bits one needs to put into a message to tell someone else the value of the element of the distribution that was picked at random.
 
@@
 Also, technically, for continuous variables the expression has a slightly different meaning: the uniform distribution in a unit volume is assigned zero entropy so $I(\rho(\xi^a))$ represents the relative number of bits to achieve the same level of precision, which can be negative.
-That is, if we take a real number from a uniform distribution between 0 and $\frac{1}{2}$ we need to lose a bit of information to get to the same precision of a unitary distribution.
+That is, if we take a real number from a uniform distribution between 0 and $\frac{1}{2}$ we need to lose a bit of information to get to the same precision of a uniform distribution over a unitary interval.
+But these technical details do not change the general intuition.
 
-Technicality aside, the above discussion, while not a formal derivation, should give a sense of what the expression signifies: it measures the \emph{additional} information that would be needed to identify the value associated to an element picked at random from a known population.
+The above discussion, while not a formal derivation, should give a sense of what the expression signifies: it measures the \emph{additional} information that would be needed to identify the value associated to an element picked at random from a known population.
+Distribution that are more spread will have a higher information entropy.
 What is also important is that it is a precisely defined quantity and that there is only one way to define it.
@@
 But mathematics only keeps track of the number, which changes, and is therefore coordinate dependent.
-In general, if $\hat{\xi}^a=\hat{\xi}^a(\xi^b)$ we can recover how the density changes by requiring that the total in each volume is the same no matter what variables are used.
+In general, if $\hat{\xi}^a=\hat{\xi}^a(\xi^b)$ is the change of variable, we can recover how the density changes by requiring that the total in each volume is the same no matter what variables are used.
 That is:
@@
 We now turn our attention to what \cite{AoPPhy1} calls an infinitesimally reducible system.
-This is a system that can be thought as made of infinitesimally small parts, called particles, and for which giving a description of the full system is equivalent to giving a full description for all infinitesimal parts.
+This is a system that can be repeatedly divided into smaller and smaller parts and for which giving a description of the full system is equivalent to giving a full description for all parts.
+We call particles the limit of this recursive subdivision.
+Note that particles here are not point-like: they have an infinitesimal size for each dimension along which the system was subdivided.
 Under this premise each particle is assigned a state from a state space $\mathcal{S}$ and the state of the whole system will be described by a distribution over those particle states.
@@
 But for states to be physically objective, they must be independent of reference frame.
-For example, if $A$ and $B$ are two reference frames and $\xi^a$ and $\hat{\xi}^b$ are the state variables one would use in the respective reference frames, we must have $s(\xi^a) = s(\hat{\xi}^b)$.
+For example, if $s$ is the state, $A$ and $B$ are two reference frames and $\xi^a$ and $\hat{\xi}^b$ are the state variables one would use in the respective reference frames, we must have $s(\xi^a) = s(\hat{\xi}^b)$.
 Therefore, since the distribution is defined on the state itself, we must have $\rho(s(\xi^a))=\rho(s(\hat{\xi}^b))$.
@@
 
-And here lies the solution to the puzzle: there is a difference between state variables, quantities used to identify a state, and coordinates, state variables that additionally define the system of reference and units.
+And here lies the solution to the puzzle: there is a difference between state variables, quantities used to identify a state, and coordinates, particular state variables that additionally define the system of reference and units.
 The density and information entropy must be invariant under arbitrary change of coordinates, not arbitrary change of state variables.
@@
 units of $1/m^3$), a normalized density for particle states will be defined over a region of possible states.
-By convention, we use $\hbar$ as the unit to describe the amount of possible states in a region.
+By convention, we use $\hbar$ as the unit to describe the number of possible states in a region.
 We have:
@@
 \end{equation}
-where we set $p=\hbar k$. We recognize $p$ as conjugate momentum and $k$ the classical analogue of the wave number.
+where we set $p=\hbar k$.
+We recognize $p$ as conjugate momentum and $k$ the classical analogue of the wave number.
+Note that, at this point, $\hbar$ is simply a constant that defines the unit for the number of states.
+It is our choice to set it to a particular value.
+That is, we do not yet know that there is a ``special number of states'' that represents a ``smallest size'' in some sense.
+That requires an extra assumption which we will discuss when looking at the classical uncertainty principle.
 
@@
 \end{equation}
-is coordinate invariant.
-In fact, creating a marginal distribution means giving a foliation (a parametrized family of surfaces that do not intersect and whose union is the whole space) and then calculating the integral as a function of the parameters.
+is coordinate invariant.\footnote{In fact, creating a marginal distribution means giving a foliation (a parametrized family of surfaces that do not intersect and whose union is the whole space) and then calculating the integral as a function of the parameters.}
 
@@
 What we are showing is that that role is not a coincidence: it is the very defining characteristic.
-\textbf{Hamiltonian mechanics is a type of statistical mechanics}, one with no probability and deterministic and reversible laws.
+\textbf{Hamiltonian mechanics is a type of statistical mechanics}, one with deterministic and reversible laws.
 
@@
 Under this light, asking whether classical mechanics is tenable coincides with asking whether we can really keep dividing our object indefinitely.
-If one is convinced that this is not realistic then one is already convinced that classical mechanics will not hold in those regimes.
-But if that approximation fails, we do not need an altogether new set of ideas: the macroscopic object is still the same.
-We simply need to change how the subdivision works.
+This approach also fails more gracefully.
+If we started with the idea of point-like objects and that idealization fails, any aggregate will also fail.
+But since we start with the macroscopic object and subdivide, the macroscopic object still remains what it is even if the idealization of infinite subdivision fails.
 
 Third, this new approach is conceptually more compatible with the other physical theories.
-We have already seen how elements of special relativity (e.g.
+We have seen how elements of special relativity (e.g.
 coordinate invariance, momentum as covector), statistical mechanics (e.g.
@@
 phase space, Poisson brackets) already emerged naturally.
-Furthermore, note that in general relativity point-like particles would be singularities that puncture space-time and would not move along a geodesics, while particles as infinitesimal parts would.
+Furthermore, note that in general relativity point-like particles would be singularities that puncture space-time and would not move along geodesics, while particles as infinitesimal parts would.
 Note that if the evolution is deterministic and reversible, it means the evolution of the system depends only on itself, which means that the system is isolated.
-That is: we have an association between isolated system and conservation of energy.\footnote{To be precise, conservation of energy is not an invariant property.
+That is, we have an association between isolated system and conservation of energy.\footnote{To be precise, conservation of energy is not an invariant property.
 The fact that a Hamiltonian exists, though, is an invariant property.} This sounds strikingly close to the first law of thermodynamics.
@@
 So we ask: what is the distribution that minimizes the spread given a certain amount of information entropy?
-If $I_0$ is the set amount of information entropy and $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$ is the spread, we can use Lagrange multipliers to find out.
+If $I$ is the set amount of information entropy and $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$ is the spread, we can use Lagrange multipliers to find out.
 \begin{align*}
 L = &\int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp \\
-&+ \lambda_1(\int \rho dqdp - 1) \\ &+ \lambda_2(- \int \rho \ln \rho \, dqdp - I_0)\\ 
+&+ \lambda_1(\int \rho dqdp - 1) \\ &+ \lambda_2(- \int \rho \ln \rho \, dqdp - I)\\ 
 \delta L = &\int \delta \rho [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho - \lambda_2 ] dqdp = 0 \\
@@
 \rho = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
-I_0 = &\ln (2\pi\sigma_q\sigma_p) + 1
+I = &\ln (2\pi\sigma_q\sigma_p) + 1
 \end{align*}
@@
 \begin{align*}
-\sigma_q\sigma_p \geq \exp (I_0 - 1) / 2 \pi 
+\sigma_q\sigma_p \geq \exp (I - 1) / 2 \pi 
 \end{align*}
-Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and its elements, and therefore it cannot shrink the distribution more than a set amount.
+Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and its elements, and therefore it cannot reduce the spread of the distribution indefinitely.
 If it did, it would start concentrating the density, meaning that multiple states would be mapped to a single state and the dynamics would not be reversible.
-In this form $I_0$ is not a fundamental constant but a parameter that depends upon the initial condition.
+In this form $I$ is not a fundamental constant but a parameter that depends upon the initial condition.
 While this is already striking, \cite{AoPPhy1} shows a simple argument to turn this into something more.
@@
 
-Now suppose that source of chaotic motion is something intrinsic to a physical system, something it cannot be avoided.
+Now suppose that source of chaotic motion is something intrinsic to a physical system, something that cannot be avoided.
 Maybe it is caused by interaction with the environment, from which the system can never be fully isolated.
@@
 Or maybe it is caused by some background process with space itself, since the vacuum is not just empty space.
-The point here is not to identify the source of this non-deterministic motion: the point is to show that assuming one exists does not sound unreasonable.
+The point here is not to identify the source of this non-deterministic motion, the point is to show that assuming one exists does not sound unreasonable.
 
-If one buys into the argument, one can posit that there exists an $I_\emptyset$ associated with this chaotic process that acts as the theoretical limit for the information entropy of our distributions.
+If one buys into the argument, one can posit that there exists an $I_0$ associated with this chaotic process that acts as the theoretical limit for the information entropy of our distributions.
 The farther the information entropy of our distribution is from that value, the more we can assume that the parts of the distribution evolve deterministically.
-The closer it is, the more our assumptions breaks down.
+The closer it is, the more our assumption breaks down.
+
+We can then decide to measure volumes in phase space in terms of this minimum spread, which means setting the value of $\hbar$ (or $h = 2 \pi \hbar$) according to the following relationship:
+\begin{align*}
+\frac{\hbar}{2} &= \frac{h}{4\pi} \equiv \frac{e^{I_0 - 1}}{2 \pi} =  \frac{e^{I_0}}{2 e \pi} \leq \sigma_q\sigma_p
+\end{align*}
  This is what \cite{AoPPhy1} refers to as the classical uncertainty principle.
@@
 
-We mentioned several time how this approach allows us to give precise physical meaning to the different mathematical objects.
-We leave in table \ref{dictionary} a summary of these connections in form of a math-to-physics dictionary.
+We mentioned several times how this approach allows us to give precise physical meaning to the different mathematical objects.
+We leave in table \ref{dictionary} a summary of these connections in the form of a math-to-physics dictionary.
 \begin{table}[h!]
@@
 		$k_i$ & Conjugate coordinate & a state variable that uses the inverse unit of the corresponding coordinate \\
-		$\hbar$ & & the unit for measuring the range of possible states within a single degree of freedom \\
+		$\hbar$ & & the unit for quantifying the number of possible states within a single degree of freedom \\
 		$p_i=\hbar k_i$ & Conjugate momentum & a state variable that together with the corresponding coordinate defines invariant ranges of possible states \\
 		$(q^i, p_i)$ & Canonical variables \newline (Canonical coordinates) & a set of variables for which the density is invariant and expressed over units of $\hbar$ per degree of freedom\\ 
-		& Canonical transformation & a change in state variables that does not change the density\\ 
+		& Canonical transformation & a change in state variables that does not change the value of the density\\ 
 	\end{tabular}
@@
 A symplectic manifold is used to define coordinate invariant densities.
-Non-canonical transformations, in fact, have non unitary Jacobians and therefore the density would change the value of the density.
+Non-canonical transformations, in fact, have non-unitary Jacobians and therefore the density would change its value.
 

### 2019-01-10 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Incorporating final comments

@@
 \title{Hamiltonian mechanics is conservation of information entropy}
-\author{Gabriele Carcassi, Christine A. Aidala}
+%\author{Gabriele Carcassi, Christine A. Aidala}
 
@@
 
-\date{\today}
+%\date{\today}
 
@@
 If $t_{in}$ is time in the inertial frame, the time $t$ in the non-inertial frame is such that  $\frac{dt}{dt_{in}} = e^{\frac{b}{m}t}$.
-Mathematical equivalence does not imply physical equivalence, which is especially problematic for \cite{Barrett2} since it focuses on the former.
+Mathematical equivalence does not imply physical equivalence, which is especially problematic for~\cite{Barrett2} since it focuses on the former; for similar arguments see~\cite{Coffey}.
 In other words, we can have the same physics with different math and the same math with different physics.
@@
 As we do not have a one-to-one map between math and physics, fixating on the mathematical structure may even confuse us.
-For example, it is known\cite{AllHamFreeParticle} that any Hamiltonian system is, at least locally, equivalent to a free particle.
+For example, it is known~\cite{AllHamFreeParticle} that any Hamiltonian system is, at least locally, equivalent to a free particle.
 That is, we can always find a canonical transformation (i.e.
 a transformation that does not change the equation of motion) such that any Hamiltonian system, in the new coordinates, has the motion of a free particle for a finite amount of time.
-As \cite{North} notes, only frame-independent relationships should be taken as fundamental.
+As~\cite{North} notes, only frame-independent relationships should be taken as fundamental.
 Does that mean that the Hamiltonian itself is not fundamental because it is not invariant under canonical transformations?
 Does it mean that all Hamiltonian systems are the same system, since there is always a set of canonical coordinates in which $H=p^2/2m$?
-It is also known\cite{AllSystemsAreHam} that for any first-order system of ordinary differential equations we can find an appropriate symplectic structure such that the system can be written as a Hamiltonian system.
+It is also known~\cite{AllSystemsAreHam} that for any first-order system of ordinary differential equations we can find an appropriate symplectic structure such that the system can be written as a Hamiltonian system.
 Does that mean that, in the end, all systems are physically equivalent, regardless of what they describe?
@@
 This leads us to the second problem we believe the previous works may have: when trying to attach meaning the thinking and intuition is still Newtonian.
-This is particularly evident in \cite{Curiel} where physical ideas provide the starting point to get to Lagrangian dynamics.
+This is particularly evident in~\cite{Curiel} where physical ideas provide the starting point to get to Lagrangian dynamics.
 The starting point is that a classical system is one that, roughly, obeys Newton's second law and therefore we have, for a free particle, $\dot{x} = v$ and $\dot{v} = 0$.
 Unfortunately this is only true in an inertial frame.
-As \cite{North} notes, both Hamiltonian and Lagrangian mechanics are coordinate independent: the equations are valid in all reference frames even non-inertial ones.
+As~\cite{North} notes, both Hamiltonian and Lagrangian mechanics are coordinate independent: the equations are valid in all reference frames even non-inertial ones.
 In fact, that is probably the most notable advantage of the frameworks, that fictitious forces are automatically handled.
@@
 It should be clear that we do not fault the authors.
-We fault ourselves, the physicists.
-Even just a century ago, there was a push in our field to find fundamental principles or laws that could serve as a starting point for the different theories.
+Even just a century ago, there was a push in physics to find fundamental principles or laws that could serve as a starting point for the different theories.
 Newtonian mechanics, thermodynamics and special relativity serve as good examples.
@@
 
-As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work\cite{AoPPhy1} that aims to identify such starting points.\footnote{We should also mention \cite{Maudlin} as another attempt to formally derive mathematical structure from physical ideas.} Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
+As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work~\cite{AoPPhy1} that aims to identify such starting points.\footnote{We should also mention~\cite{Maudlin} as another attempt to formally derive mathematical structure from physical ideas.} Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
 a symplectic structure where variables are organized in conjugate pairs).
 
-We believe this insight may be of interest to the philosophy of science community for at least two reasons.
-First because it helps to clearly characterize what Hamiltonian mechanics describes, providing a clear physical meaning to the mathematical structures used.
+This insight helps to clearly characterize what Hamiltonian mechanics describes, providing a clear physical meaning to the mathematical structures used.
 This should be at least of interest to that part of the philosophy community interested in the foundations of classical mechanics.
@@
 Before proceeding, though, let us give a quick introduction so that we have a clear understanding about what information entropy represents.
-In the physics community, for example, the concept is often vaguely referred to as ``knowledge" (and sometimes as ``lack of knowledge") which is incorrect and misleading.\footnote{We believe this characterization may be due to Jaynes\cite{Jaynes} who first introduced the concept of information theory within thermodynamics.
+In the physics community, for example, the concept is often vaguely referred to as ``knowledge" (and sometimes as ``lack of knowledge") which is incorrect and misleading.\footnote{We believe this characterization may be due to Jaynes~\cite{Jaynes} who first introduced the concept of information theory within thermodynamics.
 Note that we are only going to talk about information entropy and not thermodynamic entropy.
 We do believe there is a link between the two, but it is not the one Jaynes gives, which we believe to be unphysical.
-We leave that discussion for later works.} Information entropy was introduced by Shannon\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and should be treated as such.
+We leave that discussion for later works.} Information entropy was introduced by Shannon~\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and should be treated as such.
 What he needed was a way to measure how much information a particular channel is able to transfer and therefore a way to quantify precisely information as number of bits.
@@
 \begin{equation}
-I(\rho(\xi^a)) = \int \rho(\xi^a) \ln \left(\frac{1}{\rho(\xi^a)}\right) d\xi^1 d\xi^2 ... d\xi^n =-\int \rho(\xi^a) \ln (\rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n
+I(\rho(\xi^a)) = \int \rho(\xi^a) \ln \left(\frac{1}{\rho(\xi^a)}\right) d\xi^1 d\xi^2 ... d\xi^n =-\int \rho(\xi^a) \ln (\rho(\xi^a)) d\xi^1 d\xi^2 ... d\xi^n
 \end{equation}
@@
 The above discussion, while not a formal derivation, should give a sense of what the expression signifies: it measures the \emph{additional} information that would be needed to identify the value associated to an element picked at random from a known population.
-Distribution that are more spread will have a higher information entropy.
+Distributions that are more spread will have a higher information entropy.
 What is also important is that it is a precisely defined quantity and that there is only one way to define it.
@@
 
-We now turn our attention to what \cite{AoPPhy1} calls an infinitesimally reducible system.
+We now turn our attention to what~\cite{AoPPhy1} calls an infinitesimally reducible system.
 This is a system that can be repeatedly divided into smaller and smaller parts and for which giving a description of the full system is equivalent to giving a full description for all parts.
@@
 Under this premise each particle is assigned a state from a state space $\mathcal{S}$ and the state of the whole system will be described by a distribution over those particle states.
-The elements of our distribution are the particles and the property that characterizes each element is its state.
+The elements of our distribution are the particles, and the property that characterizes each element is its state.
 
@@
 Therefore, since the distribution is defined on the state itself, we must have $\rho(s(\xi^a))=\rho(s(\hat{\xi}^b))$.
-In this case, it seems we \emph{do} want the density to be defined on the actual space $\mathcal{S}$ and it seems it will have to be invariant under an arbitrary change of state variables.
+In this case, it seems we \emph{do} want the density to be defined on the actual space $\mathcal{S}$, and it seems it will have to be invariant under an arbitrary change of state variables.
 Accordingly, if we want densities and information entropy to be conserved over time, we first need to make sure these quantities are defined in a way that is observer independent, since otherwise they would be conserved in one frame but not in another.
@@
 Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it is not just that the density and the areas are conserved in time: it is that they are the same for all canonical transformations including change of coordinates.
-That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is precisely the evolution of (classical) distributions over a manifold that preserves information entropy and densities}.
+That is, we have an objective way to compare them at different points.} Therefore \textbf{Hamiltonian mechanics is precisely the evolution of (classical) distributions over a manifold that preserves information entropy and densities}.
 
@@
 But it is still a valid change of state variables and in fact, giving that velocity is gauge invariant, a very useful one.
-It allows to write Poisson brackets and find invariants of motion that are readily physically meaningful\cite{Holten}.
+It allows to write Poisson brackets and find invariants of motion that are readily physically meaningful~\cite{Holten}.
 
@@
 Unfortunately, mathematical structures strip out the units, the knowledge of the double role of coordinates is lost and then a different mathematical concept is used, one that has the same mathematical features of our particle state space but an altogether different meaning and construction.
-In the same vein, changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects from a cotangent bundle to a tangent bundle: it is simply relabeling the same states much like changing spatial coordinates relabels positions.
-This change of variables, though, is not in general a canonical transformation which means it does not preserve $\omega$ so the density will not correspond to the invariant one.
-But it is still a valid change of state variables and in fact, given that velocity is gauge invariant, a very useful one.
-It allows one to write Poisson brackets and find invariants of motion that are readily physically meaningful\cite{Holten}.
+We could, in many cases, use position and velocity as state variables.
+While these would not be canonical coordinates and thus not a good choice to track our densities, they are still valid state variables and in fact, given that velocity is gauge invariant, very useful ones~\cite{Holten}.
+But the space has not changed: we still have states.
+The tangent and cotangent bundle structures, then, are just tracking how conjugate momentum and velocity change differently under coordinate transformation.
 
@@
 
-The usual approach posits point-like particles, with a particular state space and equations of motion.
-Then, coincidentally, finds this setting convenient for statistical mechanics because it has nice properties.
-In retrospect, this should sound backwards.
-We believe the approach presented here is preferable, first of all, because it justifies with a single insight the state space, the evolution and why that state space is so compatible with statistical mechanics.
+This approach clarifies what we believe is a source of confusion.
+One may think that conceptually starting with parts and aggregating them to a whole is equivalent to starting from a whole and dividing it into parts.
+This is true only in the finite case.
+In the infinite case we have to start with what is finite (either the whole or the parts) and take the limit (either division or aggregation).
+We contend that the top-down approach (i.e.
+we start with a finite whole and keep subdividing) is the correct one to understand classical mechanics.
 
-Second, it maps better to what we do in classical physics.
-We study composite systems (planets, cannonballs, beads on a wire, fluids) which we sometimes pretend we can recursively divide.
-None of those objects, in fact, has a single position or momentum: only the center of mass has a single position and momentum but those are precisely the expectation values $\int q \rho(q,p) dq dp$ and $\int p \rho(q,p) dq dp$ over the distribution.
-Under this light, asking whether classical mechanics is tenable coincides with asking whether we can really keep dividing our object indefinitely.
-This approach also fails more gracefully.
-If we started with the idea of point-like objects and that idealization fails, any aggregate will also fail.
-But since we start with the macroscopic object and subdivide, the macroscopic object still remains what it is even if the idealization of infinite subdivision fails.
+First because it maps better to what we do in classical physics.
+We study composite systems (planets, cannonballs, beads on a wire, fluids) which we pretend we can recursively divide.
+None of those objects has a single position or momentum: they have an (idealized) center of mass with position and momentum given by the expectation values $\int q \rho(q,p) dq dp$ and $\int p \rho(q,p) dq dp$ of the distribution of its parts.
+A ``point particle'' is really a peaked distribution whose extent can be disregarded for the problem at hand.
 
-Third, this new approach is conceptually more compatible with the other physical theories.
+Second because it solves a set of conceptual problems: if we start with point-like objects of zero size, how can we get a finite size?
+Shouldn't we have a discrete topology?
+How could we have a differentiable structure?
+We contend that measure, topologies, metrics, differentiable structure and so on are there to remind us that the points are infinitesimal limits, keep track of how they were taken and which aggregations of those limits correspond to a finite whole.
+
+Third because it is conceptually more compatible with the other physical theories.
 We have seen how elements of special relativity (e.g.
@@
 
-The larger point is that physics is one and it cannot truly be understood piecemeal.
+The larger point, then, is that physics is one and it cannot truly be understood piecemeal.
 Nature does not separate between dynamical systems and thermodynamics, between topologies and $\sigma$-algebras, between classical systems and quantum systems.
@@
 There is no real understanding of one without the other.
-
-Having a coherent set of ideas, a coherent picture across the different areas is not only intellectually more satisfying, but it allows us to switch perspective when the intuition from one fails.
+This approach helps to resolve this by giving us a more coherent picture that spans across the different areas, which is not only intellectually more satisfying but it allows us to switch perspective when the intuition from one fails.
 As a practical example, we can show how understanding the link between Hamiltonian mechanics and information theory leads to a classical analogue for the uncertainty principle.
@@
 \end{align*}
-Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and its elements, and therefore it cannot reduce the spread of the distribution indefinitely.
+Hamiltonian mechanics is the deterministic and reversible evolution for a distribution and its elements, and therefore it cannot reduce the spread of the distribution indefinitely.
 If it did, it would start concentrating the density, meaning that multiple states would be mapped to a single state and the dynamics would not be reversible.
 In this form $I$ is not a fundamental constant but a parameter that depends upon the initial condition.
-While this is already striking, \cite{AoPPhy1} shows a simple argument to turn this into something more.
+While this is already striking,~\cite{AoPPhy1} shows a simple argument to turn this into something more.
 
-Suppose that we have a distribution where the particles move non-derministically.
+Suppose that we have a distribution where the particles move non-deterministically.
 That is, their motion is partly caused by some other agent in a way that can be characterized by random motion.
@@
 \end{align*}
- This is what \cite{AoPPhy1} refers to as the classical uncertainty principle.
+ This is what~\cite{AoPPhy1} refers to as the classical uncertainty principle.
 

### 2019-01-19 Articles/HamiltonianMechanicsIsConservationOfEntropy/titlepage.tex
Revision submitted to BJPS

--- 2018-12-07 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
@@
 
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{graphicx}
-\usepackage{amsfonts}
-\usepackage{dutchcal}
-\usepackage{braket}
-\usepackage{enumitem}
-
-\usepackage{tikz}
-\usepackage{forest}
-\usetikzlibrary{trees}
-\usetikzlibrary{calc}
-\usepackage{calculator}
-\usepackage{standalone}
-
 \begin{document}
@@
 \title{Hamiltonian mechanics is conservation of information entropy}
-\author{Gabriele Carcassi, Christine A. Aidala}
-
-%\ifjournal
-%	% As the page margins are so large, the full title does not fit
-%	\titlerunning{From physical principles to classical and quantum particle mechanics}
-%\fi
-
-\date{\today}
+\author{Gabriele Carcassi, Christine A. Aidala \\ University of Michigan}
 
@@
 
-\tableofcontents
-\newpage
+\section*{Acknowledgements}
 
-\section{Introduction}
-
-Over the past decade there has been a renewed interest in the underpinning of classical mechanics.
-Different attempts and arguments have been put forward to understand what are the most important features of the various mathematical formulations, whether they are equivalent or whether one is more fundamental than the other~\cite{North,Curiel,Barrett1,Barrett2}.
-We believe, though, that there are two general problems with these attempts.
-
-The first is that they concentrate on the mathematical structure which, unfortunately, is not enough to characterize a physical theory.
-They all note how the same physical system can be described by different mathematical frameworks.
-For example, a classical system can be characterized by position and velocity in the Lagrangian framework and by position and momentum in the Hamiltonian framework.
-They do not note, though, that the converse is also true: the same mathematical equation can represent two different physical systems.
-For example, the equation $ma + bv = 0$ in the context of Newtonian mechanics can either represent a body in an inertial frame under linear drag or a body in a non-inertial frame with no forces acting on it.
-If $t_{in}$ is time in the inertial frame, the time $t$ in the non-inertial frame is such that  $\frac{dt}{dt_{in}} = e^{\frac{b}{m}t}$.
-Mathematical equivalence does not imply physical equivalence, which is especially problematic for \cite{Barrett2} since it focuses on the former.
-In other words, we can have the same physics with different math and the same math with different physics.
-The connection between math and physics is many-to-many because mathematics captures only the syntax of a particular description without its semantics, which at that point cannot be recovered.
-
-As we do not have a one-to-one map between math and physics, fixating on the mathematical structure may even confuse us.
-For example, it is known\cite{AllHamFreeParticle} that any Hamiltonian system is, at least locally, equivalent to a free particle.
-That is, we can always find a canonical transformation (i.e.
-a transformation that does not change the equation of motion) such that any Hamiltonian system, in the new coordinates, has the motion of a free particle for a finite amount of time.
-As \cite{North} notes, only frame-independent relationships should be taken as fundamental.
-Does that mean that the Hamiltonian itself is not fundamental because it is not invariant under canonical transformations?
-Does it mean that all Hamiltonian systems are the same system, since there is always a set of canonical coordinates in which $H=p^2/2m$?
-It is also known\cite{AllSystemsAreHam} that for any first-order system of ordinary differential equations we can find an appropriate symplectic structure such that the system can be written as a Hamiltonian system.
-Does that mean that, in the end, all systems are physically equivalent, regardless of what they describe?
-The point is, we cannot know whether our findings about mathematical structures are physically significant if we do not have a precise link between the mathematical symbols and their physical meaning.
-
-This leads us to the second problem we believe the previous works may have: when trying to attach meaning the thinking and intuition is still Newtonian.
-This is particularly evident in \cite{Curiel} where physical ideas provide the starting point to get to Lagrangian dynamics.
-The starting point is that a classical system is one that, roughly, obeys Newton's second law and therefore we have, for a free particle, $\dot{x} = v$ and $\dot{v} = 0$.
-Unfortunately this is only true in an inertial frame.
-As \cite{North} notes, both Hamiltonian and Lagrangian mechanics are coordinate independent: the equations are valid in all reference frames even non-inertial ones.
-In fact, that is probably the most notable advantage of the frameworks, that fictitious forces are automatically handled.
-But none of the Newtonian concepts are coordinate independent: inertial frames, absence of forces and even conservation of energy are all frame dependent.
-This, again, leads to confusion and we may miss a key element: in the transition from Newtonian to Lagrangian/Hamiltonian mechanics we expand the reference frames in which the equations are valid (from inertial frames to all frames) at the expense of restricting the class of systems we can describe (e.g.
-dissipative forces in general require a modification of the Euler-Lagrange equation typically using the Rayleigh dissipation function).
-The point is, Newtonian concepts may prevent us from fully understanding Hamiltonian and Lagrangian mechanics.
-
-It should be clear that we do not fault the authors.
-We fault ourselves, the physicists.
-Even just a century ago, there was a push in our field to find fundamental principles or laws that could serve as a starting point for the different theories.
-Newtonian mechanics, thermodynamics and special relativity serve as good examples.
-Nowadays, unfortunately, many of the current fundamental theories, including Lagrangian, Hamiltonian, quantum mechanics and quantum field theory take as their starting point a particular mathematical structure.
-That is, we do not know what the value of a Lagrangian means or what the symplectic form represents; we just take them as given.
-Moreover, a lot of work in theoretical physics nowadays starts by exploring novel abstract mathematical structures and then trying to see whether some physical meaning or prediction can be extracted.
-We are not going to debate here how and why this shift happened.
-We merely note that it leaves the semantics of our theories ill defined.\footnote{Which is a nice way to say that we don't really know what we are talking about.}
-
-It is only natural that, given the circumstances, a philosopher would use the only thing that seems to be well defined, the math, and try to draw understanding from there.
-But here lies the fundamental problem: that math is not created by physicists to do physics, it is created by mathematicians to do math.
-The definitions of topology, differentiable manifold, symplectic manifold, Hilbert space and so on are chosen either because they are convenient to prove theorems or because they are convenient to perform calculations.\footnote{In fact, it is common in math to have two such definitions and then prove they are equivalent.} We should not expect them to be a good match for definitions that would be physically meaningful.
-In other words, there is a very good chance that looking at just the math may lead us, like horses with blinders, in a well-defined direction which is not the physically meaningful one.
-To be clear, we do not fault the mathematicians either as they are not responsible for whether their structures are used appropriately or not within physics.
-
-As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work\cite{AoPPhy1} that aims to identify such starting points.\footnote{We should also mention \cite{Maudlin} as another attempt to formally derive mathematical structure from physical ideas.} Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
-a symplectic structure where variables are organized in conjugate pairs).
-
-We believe this insight may be of interest to the philosophy of science community for at least two reasons.
-First because it helps to clearly characterize what Hamiltonian mechanics describes, providing a clear physical meaning to the mathematical structures used.
-This should be at least of interest to that part of the philosophy community interested in the foundations of classical mechanics.
-Second because information entropy has a deep connection to the foundations of thermodynamics, therefore the strong link we establish here between information entropy and Hamiltonian mechanics should be of interest to an even wider audience.
-
-We begin in section 2 by giving a brief summary of statistical distributions in general.
-These are your typical mathematical tools used to keep track of how charge is distributed in space or how a population of individuals is distributed in age.
-We briefly introduce information entropy, which measures the number of yes or no questions one has to answer to be able to identify a particular element within a distribution.
-The main issue is that the density and the information entropy for a distribution depend, in general, on the variables and their units.
-For example, if we change our units from meters to kilometers a density of $1 kg/m^3$ would change to $10^9 kg/km^3$.
-In other words, distributions are not in general invariant under change of variables.
-In section 3 we look at distributions over states.
-We argue that since states themselves are coordinate independent, the density and information entropy associated with them cannot be coordinate dependent.
-Moreover, a deterministic and reversible system must be one that preserves information entropy over time (i.e.
-information about the system at one time is equivalent to information about the system before or after).
-The requirement of coordinate-independent distributions gives us the symplectic structure (i.e.
-conjugate pairs of position and momentum) while the requirement of conservation of information entropy gives us Hamiltonian evolution.
-In section 4 we discuss the result and provide a dictionary that allows us to give precise meaning to the different mathematical symbols, addressing the issues highlighted before.
-We also show how this understanding allows us to construct a classical analogue of the uncertainty principle.
-
-\section{Distributions}
-
-Our starting point is the standard concept of a statistical distribution.
-We have a set of elements, also called population, and a set of variables that characterize some properties of those elements.
-To each element is associated a value for each variable and we want to describe how the population is distributed over the possible values of the variables.
-For example, people by age and country of residence, amount of material among different compounds, mass or charge over different positions in space.
-In all these cases, we have a total quantity $M$ that measures how many elements we have.
-This can be either a discrete quantity, like the number of people we are considering, or a continuous quantity, like the total amount of mass.
-We also have a set of $n$ variables $\xi^a$ over which the elements are distributed and a function $m(\xi^a)$ that gives us the amount for each value possible value such that:
-\begin{equation}
-\int m(\xi^a) d\xi^1 d\xi^2 ... d\xi^n = M
-\end{equation}
-The integral is used when some of the variables $\xi^a$ are continuous, and a simple sum would be used over the discrete ones. We can also define the normalized density as:
-\begin{equation}
-\rho(\xi^a) = \frac{1}{M}m(\xi^a)
-\end{equation}
-To calculate the amount in a particular region $V$ we have:
-\begin{equation}
-m_V = \int_V M \rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n
-\end{equation}
-The above expression would represent the mass or charge within a particular region, the amount of material of some selected set of compounds or the world population in a particular set of countries and age group.
-We can also define:
-\begin{equation}
-\mu(V) =\frac{m_V}{M} = \int_V \rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n \\
-\end{equation}
-which gives the fraction of the population within that region.
-
-The normalized distribution $\rho(\xi^a)$, as defined, is a statistical distribution that represents the entirety of the system being described.
-Yet, we can give it an additional meaning of probability if we run the following thought experiment: suppose we took an element at random, each element with equal chance, what is the probability $P(\xi^a \in V)$ that the values for $\xi^a$ associated with that element are within a particular region $V$?
-The answer must be $\mu(V)$ and therefore $\rho(\xi^a)$ is the probability density.
-So $\rho(\xi^a)$ has this double role as a physical statistical distribution and a conceptual probability distribution for our thought experiment.
-This second role is our link to information theory.
-
-Before proceeding, though, let us give a quick introduction so that we have a clear understanding about what information entropy represents.
-In the physics community, for example, the concept is often vaguely referred to as ``knowledge" (and sometimes as ``lack of knowledge") which is incorrect and misleading.\footnote{We believe this characterization may be due to Jaynes\cite{Jaynes} who first introduced the concept of information theory within thermodynamics.
-Note that we are only going to talk about information entropy and not thermodynamic entropy.
-We do believe there is a link between the two, but it is not the one Jaynes gives, which we believe to be unphysical.
-We leave that discussion for later works.} Information entropy was introduced by Shannon\cite{Shannon} to solve specific engineering problems and therefore, like distance or mass, information entropy is a number that measures a well defined quantity and should be treated as such.
-What he needed was a way to measure how much information a particular channel is able to transfer and therefore a way to quantify precisely information as number of bits.
-So the idea is that we have a source that sends messages to a destination, and the message is encoded in zeros and ones.
-Information entropy is the optimal average number of zeros and ones that must be used to encode the messages.
-If the source sends fewer bits per message than required by the information entropy, some information will be lost.
-If the source sends more bits per message than required by the information entropy, some information will be redundant.
-
-To understand how this works in our context, let us go back to our distribution $\rho(\xi^a)$ and refine our thought experiment.
-Suppose two individuals, Alice and Bob, have full knowledge of the same distribution.
-Suppose Alice picks an element at random and Bob needs to know the value of the variable for the element Alice picked.
-That is, Alice is the source, Bob is the destination and the message is the value of the element Alice picked from the distribution.
-How long, on average, should the message be?
-How many ones and zeros?
-How many yes or no questions must Bob ask Alice before he knows the value?
-For example, suppose there are 40 balls of which 20 are red and 20 are green.
-In that case one yes or no question (``Is it green?"), one bit of information, would be sufficient.
-Suppose there are 40 balls of which 10 red, 10 green, 10 blue and 10 yellow.
-In this case, two questions would suffice as shown in figure \ref{twoQuestions}, since two yes or no questions, two bits of information, can distinguish 4 cases.
-
-\begin{figure}\label{twoQuestions}
-	\centering
-\begin{tikzpicture}[level distance=1.5cm,
-level 1/.style={sibling distance=6cm},
-level 2/.style={sibling distance=3cm}, sloped]
-\node {Is it red or green?}
-child {node {Is it red?}
-	child {
-		node {red - 11}
-		edge from parent node[pos=0.6, align=center, above] {Yes}
-	}
-	child {
-		node {green - 10}
-		edge from parent node[pos=0.6, align=center, above] {No}
-	}
-	edge from parent node[pos=0.6, align=center, above] {Yes}
-}
-child {node {Is it blue?}
-	child {
-		node {blue - 01}
-		edge from parent node[pos=0.6, align=center, above] {Yes}
-	}
-	child {
-		node {yellow - 00}
-		edge from parent node[pos=0.6, align=center, above] {No}
-	}
-	edge from parent node[pos=0.6, align=center, above] {No}
-};
-\end{tikzpicture}
-\caption{The decision tree for the example in the text.
-The nodes of the tree represent the questions Bob can pose to Alice to identify the color and the bottom line shows the four different cases with the corresponding encodings.}
-\end{figure}
-
-In general, if $I$ is the number of questions, the number of bits, then $C = 2 ^ I$ are the number of cases one can distinguish with those questions.
-On the other hand, suppose for a particular value $\xi_0$ of a discrete variable we have $\rho(\xi_0)=1/5$, then we know that we'd select $\xi_0$ one in 5 cases.
-That is, the normalized distribution is the inverse of the number of cases: $\rho(\xi_0) = \frac{1}{C}$.
-Therefore the information needed to identify a particular case is $I(\xi^a)=\log \frac{1}{\rho(\xi^a)}$.
-If we take the average and change the base of the logarithm, we have:
-\begin{equation}
-I(\rho(\xi^a)) = \int \rho(\xi^a) \ln \left(\frac{1}{\rho(\xi^a)}\right) d\xi^1 d\xi^2 ... d\xi^n =-\int \rho(\xi^a) \ln (\rho(\xi^a) d\xi^1 d\xi^2 ... d\xi^n
-\end{equation}
-which is the information entropy of the distribution.
-That tells us the average number of bits one needs to put into a message to tell someone else the value of the element of the distribution that was picked at random.
-
-Technically, given the more convenient choice of the natural logarithm, the unit of information is the nat, which is equal to $\frac{1}{\ln 2}$ bits.
-However, in the discussion we will refer to it as measured in bits, as we believe that the majority of the readers is more comfortable with the latter unit.
-Also, technically, for continuous variables the expression has a slightly different meaning: the uniform distribution in a unit volume is assigned zero entropy so $I(\rho(\xi^a))$ represents the relative number of bits to achieve the same level of precision, which can be negative.
-That is, if we take a real number from a uniform distribution between 0 and $\frac{1}{2}$ we need to lose a bit of information to get to the same precision of a unitary distribution.
-
-Technicality aside, the above discussion, while not a formal derivation, should give a sense of what the expression signifies: it measures the \emph{additional} information that would be needed to identify the value associated to an element picked at random from a known population.
-What is also important is that it is a precisely defined quantity and that there is only one way to define it.
-It is the only expression that satisfies three simple requirements (continuity, monotonicity and additivity) that are necessary to measure information, and that is how Shannon originally derived it.
-Insofar that we want to talk about amounts of information we have only one way to do so.
-
-\subsection*{Densities and change of variables}
-
-There is one feature about densities we need to be fully aware of: in general they are not invariant under change of variables.
-Intuitively, if we change units the density changes its numerical value.
-For example, $1$ kg/m becomes $1000$ kg/km.
-Physically, we use the unit to understand that the overall quantity is the same.
-But mathematics only keeps track of the number, which changes, and is therefore coordinate dependent.
-In general, if $\hat{\xi}^a=\hat{\xi}^a(\xi^b)$ we can recover how the density changes by requiring that the total in each volume is the same no matter what variables are used.
-That is:
-\begin{align*}
-\int_V \rho(\xi^b) d\xi^1 ... d\xi^n &= \int_V \rho(\hat{\xi}^a) d\hat{\xi}^1 ... d\hat{\xi}^n \\
-&=\int_V\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\xi^1 ... d\xi^n
-\end{align*}
-\begin{equation}\label{density_transformation}
-\rho(\xi^b) = \rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|
-\end{equation}
-the density is multiplied by the absolute value of the Jacobian determinant of the transformation.
-As the Jacobian is in general a function of the variables themselves, this means that whether the density at one point is greater or smaller than at another also depends on the choice of variables.
-
-Similarly, we can see that information entropy is not invariant
-\begin{align*}
-I(\rho(\xi^b)) &=-\int \rho(\xi^b) \ln (\rho(\xi^b)) d\xi^1 ... d\xi^n \\
-&=-\int \rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| \ln \left(\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|\right) d\xi^1 ... d\xi^n \\
-&=-\int \rho(\hat{\xi}^a) \ln \left(\rho(\hat{\xi}^a) \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\|\right) d\hat{\xi}^1 ... d\hat{\xi}^n \\
-&=-\int \rho(\hat{\xi}^a) \ln (\rho(\hat{\xi}^a)) d\hat{\xi}^1 ... d\hat{\xi}^n -\int \rho(\hat{\xi}^a) \ln \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\hat{\xi}^1 ... d\hat{\xi}^n
-\end{align*}
-\begin{equation}\label{entropy_transformation}
-I(\rho(\xi^b)) =I(\rho(\hat{\xi}^a)) -\int \rho(\hat{\xi}^a) \ln \left\|\frac{\partial \hat{\xi}^a}{\partial \xi^b}\right\| d\hat{\xi}^1 ... d\hat{\xi}^n
-\end{equation}
-The reason is that we may be changing scale and therefore level of precision required to describe our value.
-Given that the change in both density and entropy is determined by the Jacobian, we conclude that \textbf{during a change of variable the information entropy of all possible distributions does not change if and only if the density remains the same at every point under the same change of variable}.
-
-To sum up, while we would like to think of a density as a number associated to a point of a space, that is technically incorrect.
-Given a manifold $\mathcal{M}$ upon which we define the distribution, the latter is \emph{not} $\rho : \mathcal{M} \to \mathbb{R}$ in general.
-That would be a scalar function, invariant under change of variables, as a point is a point no matter how it is identified.
-Instead we additionally need a set of functions $\xi^a : \mathcal{M} \to \mathbb{R}$ and only then we can define the density as $\rho : \mathbb{R}^n \to \mathbb{R}$.
-The density value is defined on the values for a set of variables, not at a point.\footnote{As the density is really a limit, we need to know how that limit is taken, and the coordinates decide that.} That is, the density is undefined without a (differentiable) set of coordinates.
-
-\subsection*{Mathematical tools for distributions}
-
-Different fields in math deal with the coordinate dependency of densities in different ways depending on their aims, which unfortunately leads to a fractured physical understanding given that one needs ideas and results from the different areas.
-In statistics and information theory one simply accepts the transformation rules and does not try to create invariant objects.
-This gives us the proper setting to study the statistical properties but it does not help us understand which mathematical elements represent objective physical entities.
-
-In probability theory one starts with three objects: a set of outcomes $\mathcal{M}$, a set of events $\sigma_\mathcal{M}$ where each event is a collection of outcomes, and a probability measure $\mu : \sigma_\mathcal{M} \rightarrow [0,1]$ that assigns a number between zero and one to each event.
-That is, we do not assign a probability density to the points of $\mathcal{M}$ but we assign a finite probability to finite regions.
-We can then define a random variable $\xi : \mathcal{M} \rightarrow \mathbb{R}$ as a real-valued function of the outcomes.
-For each random variable we can define its cumulative distribution function $F_\xi(\xi_0)=\mu(\xi<\xi_0)$ and the probability density function is its Radon-–Nikodym derivative $\rho(\xi) = \frac{dF_\xi}{d\xi}$.
-That is, the integral is the primal object while the density is derived.
-
-This approach, if applied to statistical distributions, actually makes more physical sense: what we measure is the amount of mass in a finite region and the density is the limit for a smaller and smaller one.
-The fundamental objects (outcomes, events and probability) are all defined without reference to variables.
-It also has the advantage that discrete and continuous quantities are treated in the same way.
-But in probability theory we have no notion of coordinate systems, vectors or any other geometrical objects.
-
-In differential geometry, one starts with a manifold: a set of points $\mathcal{M}$ which can be given local coordinates $\xi^i : \mathcal{M} \to \mathbb{R}$.
-One defines a tangent space $\mathsf{T}\mathcal{M}$ where vectors live and cotangent space $\mathsf{T}^*\mathcal{M}$ were linear functions of vectors live.
-Then we define $n$-forms as multi-linear functions of $n$ vectors that return the value associated with the infinitesimal parallelepiped they form.
-That is $\nu : (\mathsf{T}\mathcal{M})^n \rightarrow \mathbb{R}$.
-As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $\mu(V)=\int_V \nu$ with no reference to coordinates.
-If $e_a$ are the basis vectors associated with $\xi^a$ and $e^b$ are linear functions such that $e^b(e_a)=\delta_a^b$, then we can express each $n$-form as $\rho(\xi^a)e_1\wedge e_2 \wedge ...
-\wedge e_n$.
-The density $\rho(\xi^a)$ is the component of the $n$-form expressed in the $\xi^a$ coordinates in a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value.
-But a negative determinant means changing orientation of the region of integration, so fixing that would bring another sign change and $\rho(\xi^a)$ would not change.
-This is equivalent to saying that some coordinate systems are right handed and some are left handed.}
-
-This approach gives us a way to understand which objects are coordinate independent and which are not.
-But it has two issues: it does not allow us to easily talk about coordinate dependent operations, such as marginal distributions and information entropy, and it may use definitions that do not map well to physical concepts.
-For example, a vector is typically defined as a map between a scalar function of the manifold to another scalar function of the manifold, or as an equivalence class of trajectories.
-It is a stretch to think of fluid velocities or electric fields in those terms.\footnote{For that same reason, we do not use the differential geometry notation $\frac{\partial}{\partial \xi^a}$ and $d\xi^a$ for vector and covector basis because they do not convey the actual physical meaning of these objects as they are used in physics.
-This is another instance of the math-for-mathematician issue.}
-
-So the situation is that we have a single clear physical object we want to study, a statistical distribution, but different and somewhat disconnected mathematical frameworks to study it.
-As noted in the introduction, this should not be surprising as the frameworks are defined by mathematicians to solve their own needs.
-But it is an issue when these are used in physics without careful consideration, as each mathematical framework may only partially capture our physical understanding.
-And this is precisely why we argue that simply examining the mathematical structure used in a physical theory is not enough to gain a complete picture: mathematics can only check for self-consistency and not for completeness or correctness.
-What we will need to do, in fact, is use elements from all these frameworks as some insights are better understood in one context and some in another.
-
-\section{Coordinate-independent distributions}
-
-We now turn our attention to what \cite{AoPPhy1} calls an infinitesimally reducible system.
-This is a system that can be thought as made of infinitesimally small parts, called particles, and for which giving a description of the full system is equivalent to giving a full description for all infinitesimal parts.
-Under this premise each particle is assigned a state from a state space $\mathcal{S}$ and the state of the whole system will be described by a distribution over those particle states.
-The elements of our distribution are the particles and the property that characterizes each element is its state.
-
-Additionally, suppose that this system evolves in a way that is deterministic and reversible, that is, the state of a particle at one time is associated with one and only one state at another, future or past, time.\footnote{Note that by reversible we mean the ability to reconstruct the past, the dual of determinism, which we take to mean the ability to predict the future.
-We do not mean a time symmetric process, where the equation of motion does not change when time is reversed, or a process that can be undone by another.} In this case, all the particles and only the particles that are in the same initial state will be mapped to the same final state: the density at the initial state must be the same as the density at the final state.
-Moreover, if Alice gave Bob enough information to identify the state of an element at one time, Alice also gave Bob enough information to identify the state of the same element at a future or past time: the information entropy will remain the same over time.
-
-While the physical setting is clear enough, it introduces a puzzle.
-States will be themselves identified by quantities, which we call state variables.
-But for states to be physically objective, they must be independent of reference frame.
-For example, if $A$ and $B$ are two reference frames and $\xi^a$ and $\hat{\xi}^b$ are the state variables one would use in the respective reference frames, we must have $s(\xi^a) = s(\hat{\xi}^b)$.
-Therefore, since the distribution is defined on the state itself, we must have $\rho(s(\xi^a))=\rho(s(\hat{\xi}^b))$.
-In this case, it seems we \emph{do} want the density to be defined on the actual space $\mathcal{S}$ and it seems it will have to be invariant under an arbitrary change of state variables.
-Accordingly, if we want densities and information entropy to be conserved over time, we first need to make sure these quantities are defined in a way that is observer independent, since otherwise they would be conserved in one frame but not in another.
-But this seems in apparent contradiction with the previous section and we need to clarify that.
-
-For starters, we concede that \textbf{only invariant distributions, those for which the density and information entropy do not change under a change of coordinates, can characterize the state of a system made of infinitesimal parts and its deterministic and reversible evolution.} Since requiring information entropy to be invariant will automatically require the density to be invariant as well and vice-versa, the requirement is one and the same.
-
-Now, if our variables are discrete, all our integrals are sums and all our definitions work: everything is already invariant.
-On the other hand, if our variables are continuous (i.e.
-the state space is a manifold) we have a problem as densities and information entropy will not be invariant under an arbitrary change of state variables.
-Our state space $\mathcal{S}$, somehow, must allow us to write densities that, under coordinate transformations, change with the Jacobian (i.e.
-something we can integrate on) but also not change at all (i.e.
-something that is coordinate independent).
-How can this possibly work?
-
-Suppose $\mathcal{S}$ is a two dimensional manifold.
-Suppose we have two variables $(q,k)$ such that $k$ uses the inverse units of $q$.
-For example, if $q$ is meters then $k$ is inverse meters which means an infinitesimal area $dq dk$ will be a pure number.
-We say $q$ is a coordinate as it is a variable that defines a unit, while $k$ is the corresponding conjugate variable as it uses the inverse unit.
-Now suppose you change coordinate.
-For example, $\hat{q}$ will be kilometers and $\hat{k}$ inverse kilometers.
-We will have $d\hat{q} d\hat{k} = dq dk$.
-Therefore we will also have $\rho(q,k) = \rho(\hat{q}, \hat{k})$: the density is invariant and so is the information entropy.
-
-And here lies the solution to the puzzle: there is a difference between state variables, quantities used to identify a state, and coordinates, state variables that additionally define the system of reference and units.
-The density and information entropy must be invariant under arbitrary change of coordinates, not arbitrary change of state variables.
-The insight is that one choice of coordinate can influence more then one state variable.
-But how many can it influence?
-
-It turns out that if we have a set of variables fixed by a single choice of unit, a single coordinate, then a pair $(q, k)$ is the only possible way.
-Let $q$ be the coordinate, the variable that defines the unit.
-Suppose we have $n$ variables $\xi^a$ such that $\xi^1 = q$.
-The vector $d\xi^a$ representing the infinitesimal displacement has $n$-components.
-If we change unit $\hat{q}=\hat{q}(q)$ we have two constraints: the unit change and density invariance, which means the Jacobian must be unitary.
-These must be enough to determine how the vector components change.
-If not, we would need to specify another relationship between the old $\xi^a$ and the new variables $\hat{\xi}^b$ which means it is not true that the units of all variables are fixed by a single choice of unit for $q$.
-As the linear system for $d\hat{\xi}^b$ must be solvable, the number of vector components must be equal to the number of constraints and therefore there are exactly two variables $\xi^a = (q, k)$.
-Then $k$ is the conjungate variable of $q$.
-There is no other choice.
-
-To sum up, the density and information entropy will be invariant under coordinate transformations but will not be, in general, invariant under all state variable transformations.
-Physically, the coordinates play a double role.
-They define the reference frame and also help to identify particle states.
-It is this double role that gives them a special character because a change in $q$ must induce a change in $k$ as their units and the unit of $\rho$ are not independent.
-This is the key physical insight that will lead us to Hamiltonian mechanics.
-Note that symplectic geometry strips away the units and does not elevate densities as primary objects.
-It is no wonder this insight gets lost.
-Which goes back to our original point: we need to know what the mathematical structures represent physically and that cannot be deduced from the math itself.
-
-\subsection*{Hamiltonian mechanics for one degree of freedom}
-
-We are now ready to see how conservation of information entropy leads to the phase space of Hamiltonian mechanics.
-For now, we assume our invariant distribution is characterized by only one coordinate which means two state variables $(q, k)$.
-While we have discussed units for the state variables, we still need to choose the unit for the density itself.
-In the same way a normalized density in space is a pure number over a volume (e.g.
-units of $1/m^3$), a normalized density for particle states will be defined over a region of possible states.
-By convention, we use $\hbar$ as the unit to describe the amount of possible states in a region.
-We have:
-\begin{equation}
-m_V = M \mu(V) =\int_V m(q, k) \hbar dq dk = \int_V M \rho(q, p) dq dp
-\end{equation}
-where we set $p=\hbar k$. We recognize $p$ as conjugate momentum and $k$ the classical analogue of the wave number.
-
-Now suppose we change variables. In general the Jacobian will be given by:
-\begin{equation}
-\label{Poisson}
-\begin{aligned}
-|J| &= \left| \begin{matrix}
-\dfrac{\partial \hat{q}}{\partial q} & \dfrac{\partial \hat{q}}{\partial p} \\[2.2ex]
-\dfrac{\partial \hat{p}}{\partial q} & \dfrac{\partial \hat{p}}{\partial p} \end{matrix} \right| = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} - \frac{\partial \hat{p}}{\partial q} \frac{\partial \hat{q}}{\partial p} &= \{\hat{q}, \hat{p}\}
-\end{aligned}
-\end{equation}
-which we recognize to be the Poisson bracket. If the change of variables is simply a change of coordinates, we have:
-\begin{equation}
-\label{coordinate_change}
-\begin{aligned}
-\hat{q} &= \hat{q}(q) \\
-\{\hat{q}, \hat{p}\} &= 1 = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} \\
-\dfrac{\partial \hat{p}}{\partial p} &= \frac{\partial \hat{q}}{\partial q} ^{-1} = \frac{\partial q}{\partial \hat{q}} \\
-\hat{p} &= \frac{\partial q}{\partial \hat{q}} p
-\end{aligned}
-\end{equation}
-We find that conjugate momentum transforms as a covariant component of a covector and that the density and information entropy are invariant:
-\begin{equation}
-\label{density_invariance}
-\begin{aligned}
-\rho(q,p) &= \rho(\hat{q}, \hat{p}) \\
-I(\rho(q,p)) &= I(\rho(\hat{q},\hat{p}))
-\end{aligned}
-\end{equation}
-Let us call $(q,p)$ canonical variables if the density is equal to the invariant density.
-Then two other variables such that $\{\hat{q}, \hat{p}\}=1$ are also canonical since the Jacobian determinant will be unitary.
-
-Note that we have our cake, the density transforms like a Jacobian, and we can eat it too, the Jacobian is one and therefore the density and the information entropy are invariant.
-This is exactly what we needed and we have also given a precise meaning to all the pieces: the wave number $k$ is the additional state variable that each coordinate must introduce to give invariant densities, $\hbar$ is the unit we use to count the possible states over a region which is the inverse of the unit of the invariant density, canonical variables are those that define the density in the proper unit, a canonical transformation is one that leaves the density invariant and each coordinate transformation is a canonical transformation because conjugate momentum changes like a covector.
-It is striking how much we were able to get with such a simple starting point of coordinate-independent distributions.
-
-Now it is time to find the equations of motion.
-We can think of the evolution as a vector field on $\mathcal{S}$ of components $S = (S^q, S^p) = (\frac{dq}{dt}, \frac{dp}{dt})$ that gives the direction in which states move in time.
-After an infinitesimal time step, we would have:
-\begin{equation}
-\begin{aligned}
-q(t+dt) &= q(t) + \frac{dq}{dt} dt = q(t) + S^q dt \\
-p(t+dt) &= p(t) + \frac{dp}{dt} dt = p(t) + S^p dt
-\end{aligned}
-\end{equation}
-The Jacobian for time evolution will be
-\begin{equation}
-\label{Jacobian_evolution}
-\begin{aligned}
-|J| &= \left| \begin{matrix}
-1 + \dfrac{\partial S^q}{\partial q}dt & \dfrac{\partial S^q}{\partial p} dt \\[2.2ex]
-\dfrac{\partial S^p}{\partial q}  dt & 1 + \dfrac{\partial S^p}{\partial p} dt \end{matrix} \right| \\
-&= 1 + \left[ \dfrac{\partial S^q}{\partial q} + \dfrac{\partial S^p}{\partial p} \right]dt + O(dt^2)\\
-&= 1 + div(S)dt + O(dt^2)\\
-\end{aligned}
-\end{equation}
-
-If we want the time evolution to transport  densities and to conserve information entropy, the Jacobian must be unitary and therefore $S$ is divergence-free and admits a potential. We have
-\begin{equation}
-\label{Potential_Hamilton}
-\begin{aligned}
-S &= \left(\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q}\right) \\
-\frac{dq}{dt} &= \frac{\partial H}{\partial p}  \\
-\frac{dp}{dt} &= - \frac{\partial H}{\partial q}  \\
-\end{aligned}
-\end{equation}
-and we recognize Hamilton's equations.
-Note that the reverse is true as well: any Hamiltonian evolution will transport densities and conserve information entropy due to Liouville's theorem.\footnote{Again, it is not just that the density and the areas are conserved in time: it is that they are the same for all canonical transformations including change of coordinates.
-That is, we have an objective way to compare them.} Therefore \textbf{Hamiltonian mechanics is precisely the evolution of (classical) distributions over a manifold that preserves information entropy and densities}.
-
-We would like to stress how, once the simple premise of invariant distributions under deterministic and reversible evolution is accepted, all elements of Hamiltonian mechanics are necessary and have a straightforward physical meaning.
-Not only that: because of the equivalence we know that everything in Hamiltonian mechanics can be understood with just that simple premise.
-And the premise is not about cotangent bundles of Lie groups of quasi-q spaces on an infinite-dimensional lattice.
-It is just the simple realization that if we require distributions to have a physically objective reality they must be coordinate independent.
-The premise is much simpler than the math employed.
-We can reason intuitively on the premise.
-We can think about what it means and when it is applicable.
-We know what we are talking about.
-
-Note how we have not talked about mass, acceleration or forces.
-None of those Newtonian concepts appeared.
-And since we were able to get phase space and the equations of motion without them, it means they cannot play any significant role.
-Note how all the concepts we used, states and parts of a system, are automatically coordinate invariant.
-Note how it is clear what type of time evolution we are \emph{not} describing: those that are non-deterministic or non-reversible.
-Those that would concentrate or spread our distributions, those that would change the amount of information to describe a particle state at a set level of precision are ruled out.
-
-Note how each mathematical object found its natural place.
-Note how we had to cross the boundaries of the different mathematical frameworks to find meaning as the current mathematical frameworks we have do not lend themselves to talk about issues of units and coordinate-invariance of information entropy.
-Note, though, that our reasoning is logically necessary and therefore, in principle, we should be able to create a physically motivated mathematical framework that would capture these ideas.
-We just cannot expect mathematicians to magically do it for us.
-
-\subsection*{Distributions over multiple degrees of freedom}
-
-To understand what happens with multiple degrees of freedom, let us first look at change of variables applied to marginal distributions.
-Suppose we have a distribution $\rho(\xi^a)$ that depends on many coordinates, each with its unit.
-If the density is coordinate invariant, then it will be so even if we change only one coordinate, say $\xi^1$.
-As we saw before, then we can find the conjugate variable $\xi^2$ that changes in the opposite way.
-We can calculate the marginal distribution by integrating over those two quantities.
-\begin{equation}
-\rho(\xi^b) = \int \rho(\xi^a) d\xi^1 d\xi^2
-\end{equation}
-where $\xi^b$ are the remaining coordinates.
-If the variables $\xi^b$ are truly independent from $\xi^1$ and $\xi^2$, then a change of coordinates in the first will not affect the others and vice-versa.
-This means $\rho(\xi^b)$ must be coordinate invariant.
-In terms of information theory, invariance of the marginal densities is equivalent to being able to assign a coordinate invariant information entropy to each of the marginal distributions.
-
-We can imagine proceeding iteratively, integrating away one pair at a time.
-We will find that all our variables are paired $(q^i, p_i)$  with the first ones defining the units and the second ones defining the invariant areas.
-We call each pair an independent degree of freedom as it constitutes an independent choice of unit.
-We recognize $\mathcal{Q}$, the manifold charted by $q^i$, as the (poorly named) configuration space.
-We are going to call it coordinate space as it includes all variables and only the variables that define the units and the reference frame.
-The state space for the particles, then, is a set of coordinates plus an equal number of conjugate variables that vary like covector components.
-
-The idea, then, is that a distribution is truly coordinate-independent only when the density and information entropy for all the possible marginal distributions on any combination of degrees of freedom are invariant, not just for the joint distribution.
-
-%Now, the fact that we can use a tangent bundle to track our particle states does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number.
-The state space is just a manifold, a set of points identified by $n$ continuous variable, and the structure it has comes from how units transform, not because the variables identify separate objects.
-Unfortunately mathematical structures strip out the units, the knowledge of the double role of coordinates is lost and then a different mathematical concept is used, one that has the same mathematical features of our particle state space but an altogether different meaning and construction.
-Changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects from a cotangent bundle to a tangent bundle: it is relabeling the same states much like changing spatial coordinates relabels positions.
-This change of variables, though, is not in general a canonical transformation which means the density will not correspond to the invariant one.
-But it is still a valid change of state variables and in fact, giving that velocity is gauge invariant, a very useful one.
-It allows to write Poisson brackets and find invariants of motion that are readily physically meaningful\cite{Holten}.
-
-%We can summ
-
-% In fact, if $(q^1, p_1)$ and $(q^2, p_2)$ are two degrees of freedom and $\rho_1(q^1, p_1)$ and $\rho_2(q^2, p_2)$ are the respective marginal distributions, we have now four ways of characterizing that they are independent.
-%These characterizations are equivalent because they repre
-
-%This means that relationships between the entropy of the total and marginal distributions are preserved.
-For example, saying that the marginal distributions $\rho_1(q^1, p_1)$ and $\rho_2(q^2, p_2)$ are independent is equivalent to saying that the entropy of the joint distribution is equal to the sum of the entropy of the marginals: $I(\rho(q^1,q^2, p_1, p_2)) = I(\rho_1(q^1, p_1)) +I(\rho_2(q^2, p_2))$.\footnote{This is also equivalent to saying that the joint distribution is the product of the marginals $\rho(q^1,q^2, p_1, p_2) = \rho_1(q^1, p_1)\rho_2(q^2, p_2)$.} That is, giving both values is equivalent to giving one and then the other.
-This means we can think of these relationships and quantities as real physical properties assigned to the degree of freedom and not just to the coordinates.
-
-%\begin{description}
-%	\item \emph{Coordinate independence}. The unit of $q^1$ can be chosen independently of $q^2$.
-%	\item \emph{State count}. If $dq^1 dp_1$ and $dq^2 dp_2$ represent how many possible states the count of states in each degree of freedom then the total count across is the product 
-%	\item \emph{Statistical independence}. The joint distribution is the product of the marginals: $\rho(q^1,q^2, p_1, p_2) = \rho_1(q^1, p_1)\rho_2(q^2, p_2)$.
-%	\item \emph{Information independence}.
-The information entropy of the joint distribution is the sum of the information entropy of the marginals: $I(\rho(q^1,q^2, p_1, p_2)) = I(\rho_1(q^1, p_1)) +I(\rho_2(q^2, p_2))$.
-Intuitively this means that giving the value of all state variables is equivalent to giving the values of both degree of freedom one after the other.
-%\end{description}
-%These characterizations are equivalent: you can't have one without the others.
-
-\subsection*{The geometry of phase space}
-
-Now that we have an intuitive sense of the objects we want to describe physically, we can turn to differential geometry to derive the geometrical structure of the state space and then the equations of motion.
-What we want to capture is that we can integrate densities over two dimensional sub-manifolds in a way that is coordinate invariant.
-That is, we need to have a two-form $\omega$ such that:
-\begin{equation}
-\rho_\Sigma = \int_\Sigma \rho(\xi^a) \omega(d\Sigma)
-\end{equation}
-is coordinate invariant.
-In fact, creating a marginal distribution means giving a foliation (a parametrized family of surfaces that do not intersect and whose union is the whole space) and then calculating the integral as a function of the parameters.
-
-The requirement that $\omega$ is a two-form, i.e.
-linear and anti-symmetric, comes from the fact that it needs to represent areas.
-That is, it needs to be linear because the area of a parallelogram is a linear function of its sides.
-And if $v^1$ and $v^2$ are vectors and $\omega(v^1, v^2)$ represents the area, then rotating by 90 degrees will give the same area.
-That is, $\omega(v^1, v^2) = \omega(v^2, -v^1) = -\omega(v^2, v^1)$.
-Therefore it must be anti-symmetric.
-It should also be non-degenerate, i.e.
-there is no direction for which $\omega$ is always zero or we would not be able to define a volume.
-
-We also want the two-form to be closed, that is the integral
-\begin{equation}
-\int_\Sigma \omega(d\Sigma)
-\end{equation}
-is zero if $\Sigma$ is a closed surface.
-To understand why, note that it corresponds to the case where $\rho(\xi^a)=1$.
-If the surface is a parallelepiped in some variables, each face that gives a positive contribution will have a parallel face that would give an equal and opposite contribution.
-As the closed integral needs to be zero for all parallelepipeds, it will also be zero for an arbitrary closed surface.
-
-Under a change of variables, the two-form $\omega$ must be invariant.
-That is, the particle state space $(\mathcal{S}, \omega)$ is a symplectic manifold.
-And, as any symplectic manifold, we can find Darboux variables $q^i, p_j$ and express the two-form as
-\begin{equation}
-\label{Symplectic}
-\begin{aligned}
-\omega &= dq^i \wedge dp_i = d\xi^a\omega_{ab}d\xi^b \\
-\omega_{ab} &=  \left[
-\begin{array}{cc}
-0 & 1 \\
--1 & 0 \\
-\end{array}
-\right] \otimes I_n =
-\left[
-\begin{array}{cc}
-0 & I_n \\
--I_n & 0 \\
-\end{array}
-\right]
-\end{aligned}
-\end{equation}
-
-The components $\omega_{ab}$ let us understand what geometrical features are captured by $\omega$.
-It corresponds to a two dimensional vector product within each degree of freedom while it corresponds to a scalar product across degrees of freedom.
-As $\omega$ is preserved during coordinate transformations, the area within each degree of freedom is coordinate independent meaning areas of phase space are invariant and so are the marginal densities and information entropies.
-Across degrees of freedom, instead, we are preserving the angle meaning two perpendicular degrees of freedom remain perpendicular.
-But here perpendicularity means independence, so whether two degrees of freedom are independent is now invariant as well.
-As these concepts are coordinate invariant, they correspond to actual physical objects that all observers can agree on.
-They are part of the objective physical reality being described.
-
-Now, the fact that we can use symplectic manifolds, or tangent bundles, to track our particle states does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number, defined on the coordinate space.
-The state space comes as one whole manifold, a set of points identified by $n$ continuous variables $\xi^a$, and the structure it has comes from how units transform, not because the variables identify separate objects.
-Unfortunately, mathematical structures strip out the units, the knowledge of the double role of coordinates is lost and then a different mathematical concept is used, one that has the same mathematical features of our particle state space but an altogether different meaning and construction.
-In the same vein, changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects from a cotangent bundle to a tangent bundle: it is simply relabeling the same states much like changing spatial coordinates relabels positions.
-This change of variables, though, is not in general a canonical transformation which means it does not preserve $\omega$ so the density will not correspond to the invariant one.
-But it is still a valid change of state variables and in fact, given that velocity is gauge invariant, a very useful one.
-It allows one to write Poisson brackets and find invariants of motion that are readily physically meaningful\cite{Holten}.
-
-\subsection*{Hamiltonian mechanics for multiple degrees of freedom}
-
-As deterministic and reversible evolution will conserve densities and information entropies, it will conserve $\omega$.
-That is, $\omega(v, w) = \omega(v',w')$ where $v'$ and $w'$ are the vectors evolved in time.
-In math terms, deterministic and reversible evolution for an invariant distribution is a symplectomorphism.
-As before, $S = \left(\frac{d\xi^a}{dt}\right)$ is the vector field corresponding to the evolution.
-The vector components change according to the Jacobian.
-\begin{equation}
-	v'^a = \partial_b \xi^a(t+dt) v^b = (\delta^a_b + \partial_b S^a dt) v^b
-\end{equation}
-Therefore we have:
-\begin{align*}
-v^{a} \omega_{ab} w^{b} &= v'^{a} \omega_{ab} w'^{b}  \\
-&= (v^{a} + \partial_{c} S^{a} v^{c} dt) \omega_{ab} ( w^{b} + \partial_{d} S^{b} w^{d} dt) \\
-&= v^{a} \omega_{ab} w^{b} + (\partial_{c} S^{a} v^{c} \omega_{ab} w^{b} + v^{a} \omega_{ab} \partial_{d} S^{b} w^{d}) dt \\ &+ O(dt^2)
-\end{align*}
-If we set $S_{b} \equiv S^{a} \omega_{ab}$, we have:
-\begin{equation}
-\begin{aligned}
-v^{c} w^{b} \partial_{c} S_{b} &- v^{a} w^{d} \partial_{d} S_{a} = 0\\
-\partial_{a} S_{b} - \partial_{b} S_{a} &= curl(S_{a}) = 0 \\
-S^{b} \omega_{ba} = S_{a} &= \partial_{a}H \\
-\frac{dq^i}{dt} &= \frac{\partial H}{\partial p_i}  \\
-\frac{dp_i}{dt} &= - \frac{\partial H}{\partial q^i}
-\end{aligned}
-\end{equation}
-
-We recognize Hamilton's equations for multiple degrees of freedom.
-Again, we stress how the symplectic structure was recovered by requiring invariant distributions which in turn requires invariant integration over independent degrees of freedom.
-Note how the requirement was given outside of the symplectic structure itself and therefore it cannot be recovered simply from that.
-
-\section{Discussion}
-
-Hamiltonian mechanics, through Liouville's theorem, has always had an important role in statistical mechanics.
-What we are showing is that that role is not a coincidence: it is the very defining characteristic.
-\textbf{Hamiltonian mechanics is a type of statistical mechanics}, one with no probability and deterministic and reversible laws.
-
-The usual approach posits point-like particles, with a particular state space and equations of motion.
-Then, coincidentally, finds this setting convenient for statistical mechanics because it has nice properties.
-In retrospect, this should sound backwards.
-We believe the approach presented here is preferable, first of all, because it justifies with a single insight the state space, the evolution and why that state space is so compatible with statistical mechanics.
-
-Second, it maps better to what we do in classical physics.
-We study composite systems (planets, cannonballs, beads on a wire, fluids) which we sometimes pretend we can recursively divide.
-None of those objects, in fact, has a single position or momentum: only the center of mass has a single position and momentum but those are precisely the expectation values $\int q \rho(q,p) dq dp$ and $\int p \rho(q,p) dq dp$ over the distribution.
-Under this light, asking whether classical mechanics is tenable coincides with asking whether we can really keep dividing our object indefinitely.
-If one is convinced that this is not realistic then one is already convinced that classical mechanics will not hold in those regimes.
-But if that approximation fails, we do not need an altogether new set of ideas: the macroscopic object is still the same.
-We simply need to change how the subdivision works.
-
-Third, this new approach is conceptually more compatible with the other physical theories.
-We have already seen how elements of special relativity (e.g.
-coordinate invariance, momentum as covector), statistical mechanics (e.g.
-joint and marginal distributions), quantum mechanics (e.g.
-wave number) and classical Hamiltonian mechanics (e.g.
-phase space, Poisson brackets) already emerged naturally.
-Furthermore, note that in general relativity point-like particles would be singularities that puncture space-time and would not move along a geodesics, while particles as infinitesimal parts would.
-Note that if the evolution is deterministic and reversible, it means the evolution of the system depends only on itself, which means that the system is isolated.
-That is: we have an association between isolated system and conservation of energy.\footnote{To be precise, conservation of energy is not an invariant property.
-The fact that a Hamiltonian exists, though, is an invariant property.} This sounds strikingly close to the first law of thermodynamics.
-
-The larger point is that physics is one and it cannot truly be understood piecemeal.
-Nature does not separate between dynamical systems and thermodynamics, between topologies and $\sigma$-algebras, between classical systems and quantum systems.
-The boundaries between all these disciplines are as much unfortunate as they are artificial.
-There is no real understanding of one without the other.
-
-Having a coherent set of ideas, a coherent picture across the different areas is not only intellectually more satisfying, but it allows us to switch perspective when the intuition from one fails.
-As a practical example, we can show how understanding the link between Hamiltonian mechanics and information theory leads to a classical analogue for the uncertainty principle.
-
-\subsection*{Classical uncertainty principle}
-
-As we saw, Hamiltonian dynamics not only conserves the energy of individual particles, but it also conserves the information entropy of the distribution as a whole.
-Intuitively, we can see that there is some link between the spread of a distribution and the number of bits required to identify an element.
-For example, if we have two uniform distributions with different ranges, then Alice will need to give Bob more information to identify the value from the bigger range.
-So we ask: what is the distribution that minimizes the spread given a certain amount of information entropy?
-If $I_0$ is the set amount of information entropy and $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp$ is the spread, we can use Lagrange multipliers to find out.
-\begin{align*}
-L = &\int (q-\mu_q)^2 \rho \, dqdp \int (p-\mu_p)^2 \rho \, dqdp \\
-&+ \lambda_1(\int \rho dqdp - 1) \\ &+ \lambda_2(- \int \rho \ln \rho \, dqdp - I_0)\\ 
-\delta L = &\int \delta \rho [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho - \lambda_2 ] dqdp = 0 \\
-\lambda_2 \ln \rho = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
-\rho = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}\\
-\end{align*}
-We solve the multipliers and have:
-\begin{align*}
-\rho = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
-I_0 = &\ln (2\pi\sigma_q\sigma_p) + 1
-\end{align*}
-The distribution that minimizes the spread, then, is the product of two independent Gaussians.
-As the entropy is conserved during Hamiltonian evolution the product $\sigma_q^2 \sigma_p^2$ can never be less than the one given by the Gaussian distribution of the same entropy.
-We have:
-\begin{align*}
-\sigma_q\sigma_p \geq \exp (I_0 - 1) / 2 \pi 
-\end{align*}
-Intuitively, Hamiltonian mechanics is deterministic and reversible evolution for a distribution and its elements, and therefore it cannot shrink the distribution more than a set amount.
-If it did, it would start concentrating the density, meaning that multiple states would be mapped to a single state and the dynamics would not be reversible.
-In this form $I_0$ is not a fundamental constant but a parameter that depends upon the initial condition.
-While this is already striking, \cite{AoPPhy1} shows a simple argument to turn this into something more.
-
-Suppose that we have a distribution where the particles move non-derministically.
-That is, their motion is partly caused by some other agent in a way that can be characterized by random motion.
-Now suppose our distribution is at an equilibrium with that random motion.
-Then the particles change state randomly but the overall distribution remains the same.
-That is, the random motion simply reshuffles the elements.
-In this case the minimum achievable for the information entropy of the distribution depends on the random motion itself.
-The stronger it is, the more spread the distribution has to be, the higher the minimum threshold for the information entropy.
-
-Now suppose that source of chaotic motion is something intrinsic to a physical system, something it cannot be avoided.
-Maybe it is caused by interaction with the environment, from which the system can never be fully isolated.
-Or maybe it is caused by some internal motion, which can never be fully accessible.
-Or maybe it is caused by some background process with space itself, since the vacuum is not just empty space.
-The point here is not to identify the source of this non-deterministic motion: the point is to show that assuming one exists does not sound unreasonable.
-
-If one buys into the argument, one can posit that there exists an $I_\emptyset$ associated with this chaotic process that acts as the theoretical limit for the information entropy of our distributions.
-The farther the information entropy of our distribution is from that value, the more we can assume that the parts of the distribution evolve deterministically.
-The closer it is, the more our assumptions breaks down.
-This is what \cite{AoPPhy1} refers to as the classical uncertainty principle.
-
-\subsection*{A dictionary}
-
-We mentioned several time how this approach allows us to give precise physical meaning to the different mathematical objects.
-We leave in table \ref{dictionary} a summary of these connections in form of a math-to-physics dictionary.
-\begin{table}[h!]
-	\centering
-	%	\begin{tabular}{p{0.2\textwidth} p{0.1\textwidth} p{0.1\textwidth} p{0.5\textwidth}}
-	\begin{tabular}{c p{0.3\textwidth} p{0.5\textwidth} }
-		& Name & Meaning\\ 
-		\hline 
-		& Classical particle & an infinitesimal part of the system (i.e. the limit of recursive division) \\ 
-		$\mathcal{S} =\mathsf{T}^*\mathcal{Q}$ & Phase space \newline (Cotangent bundle) & the set of all possible states for particles upon which invariant distributions can be defined \\
-		$\xi^a$ & State variables \newline (Unified coordinates) & the set of variables needed to identify the state of a particle \\ 
-		$\rho(\xi^a)$ & Density & the amount of material for a given particle state (i.e. in the limit of the recursive division)\\ 
-		$I(\rho(\xi^a))$ & Information entropy & the number of bits to identify an element of the distribution to a unitary level of precision\\ 
-		$q^i$ & Coordinate & a state variable that also defines a unit \\
-		$\mathcal{Q}$ & Coordinate space \newline (Configuration space) & the space charted by all coordinates \\
-		$k_i$ & Conjugate coordinate & a state variable that uses the inverse unit of the corresponding coordinate \\
-		$\hbar$ & & the unit for measuring the range of possible states within a single degree of freedom \\
-		$p_i=\hbar k_i$ & Conjugate momentum & a state variable that together with the corresponding coordinate defines invariant ranges of possible states \\
-		$(q^i, p_i)$ & Canonical variables \newline (Canonical coordinates) & a set of variables for which the density is invariant and expressed over units of $\hbar$ per degree of freedom\\ 
-		& Canonical transformation & a change in state variables that does not change the density\\ 
-	\end{tabular}
-	\caption{Dictionary between mathematical and physical objects.}
-	\label{dictionary}
-\end{table}
-
-We can also better understand what physical requirement each mathematical structure is used for.
-Manifolds are used to represent a set of physical objects that can be identified by $n$ continuous variables.
-A differentiable structure on a manifold is used to define densities.
-Non-differentiable transformations, in fact, do not define a Jacobian and therefore cannot express the density in the new variables.
-A symplectic manifold is used to define coordinate invariant densities.
-Non-canonical transformations, in fact, have non unitary Jacobians and therefore the density would change the value of the density.
-
-\section{Conclusion}
-
-In this paper we have shown how classical Hamiltonian mechanics coincides with conservation of information entropy.
-The structure of phase space is required to be able to define densities and information entropy in a way that is coordinate invariant, and therefore physically meaningful.
-The equations of motion and the existence of the Hamiltonian are required to conserve densities and information entropy over time.
-This clarifies the link between Hamiltonian mechanics, statistical mechanics and deterministic and reversible motion.
-
-In the narrower context, it allows us to understand what physical object is represented by each mathematical one.
-In the broader context, it allows us to make deeper connections among different areas of math and physics, and pushes us to rethink our starting points.
-In particular, it tells us that classical particles should not be considered point-like objects (upon which entropy and densities cannot be defined), but the limit of recursive subdivisions (i.e.
-infinitesimal cells of phase space).
-
-%These results represent one set of insights from our larger project, Assumptions of Physics, that aims to rederive the standard laws from a handful of physical starting points.
-
-This approach may point to a reorganization of fundamental physics on a more physically rigorous conceptual footing and more meaningful mathematical grounds which we believe is not only possible, but long overdue.
-
-\bibliographystyle{alpha}
-
-\bibliography{bibliography}{}
+We would like to thank Josh Hunt, David J.
+Baker and Gordon Belot for feedback.
+We would especially like to thank Mario Hubert: without his inspiration and guidance this work would have not been possible.
+This article is part of a larger project, Assumptions of Physics, that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.\footnote{See http://assumptionsofphysics.org/book for further details}
 

### 2019-01-28 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Adapting to FoP style
Updating accepted reference
Rewrote the introduction

@@
+\newif\ifarchive
+\archivetrue
+%\archivefalse
+
+\ifarchive
 	\documentclass[letterpaper]{article}
+\else
+	\documentclass[smallextended]{svjour3}
+\fi
 
 %% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
+%\usepackage[english]{babel}
+%\usepackage[utf8x]{inputenc}
+%\usepackage[T1]{fontenc}
 
 %% Sets page size and margins
-\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}
+%\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}
 
-% Temporary to remove "submitted to"
-\makeatletter
-\def\ps@pprintTitle{%
-	\let\@oddhead\@empty
-	\let\@evenhead\@empty
-	\def\@oddfoot{}%
-	\let\@evenfoot\@oddfoot}
-\makeatother
+%\usepackage{lineno,hyperref}
 
-\usepackage{lineno,hyperref}
-
+\ifarchive
+% Theorem definitions using amsthm
 \usepackage{amsmath, amsthm, amsfonts}
-\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}
 
@@
 
+\else
+
+\usepackage{amsmath, amsfonts}
+
+\numberwithin{equation}{section}
+\spnewtheorem{assump}{Assumption}{\bf}{\it}
+\renewcommand{\theassump}{\Roman{assump}}
+\spnewtheorem{prop}[equation]{Proposition}{\bf}{\rm}
+\spnewtheorem{thrm}[equation]{Theorem}{\bf}{\it}
+\spnewtheorem{defn}[equation]{Definition}{\bf}{\it}
+\spnewtheorem{axiom}[equation]{Axiom}{\bf}{\it}
+\spnewtheorem*{principle*}{Principle of scientific objectivity}{\bf}{\it}
+
+\newenvironment{rationale}{\emph{Rationale}.}{\hfill\(\qed\)}
+\newenvironment{justification}{\emph{Justification}.}{\hfill\(\qed\)}
+\renewenvironment{proof}{\emph{Proof}.}{\hfill\(\qed\)}
+\fi
+
+\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}
+
 % Remove line spaces between items of enumerate and itemize
@@
 % Allows to create negation symbols
+\let\vec\undefined
 \usepackage{MnSymbol}
@@
 \title{Towards a general mathematical theory of experimental science}
+
+\ifarchive
 \author{Gabriele Carcassi, Christine A. Aidala \\ University of Michigan}
-
 \date{\today}
+\else
+\author{Gabriele Carcassi \and Christine A. Aidala}
+\institute{G. Carcassi, C. A. Aidala \at
+	Physics Department, University of Michigan, Ann Arbor, MI 48109 \\
+	\email{carcassi@umich.edu}
+}
+\date{Received: date / Accepted: date}
+\fi
 
@@
 \begin{abstract}
-	In this article we lay the groundwork for a general mathematical theory of experimental science.
-The starting point will be the notion of verifiable statements, those assertions that can be shown to be true with an experimental test.
-We study the algebra of such objects and show how it is closed only under finite conjunction and countable disjunction.
-With simple constructions, we show that the set of possible cases distinguishable by verifiable statements is equipped with a natural Kolmogorov and second countable topology and a natural $\sigma$-algebra.
+	We lay the groundwork for a formal theory that studies scientific theories, which, as physics provides the foundation for all sciences, we feel is a neglected responsibility of the physics community.
+We define a scientific theory as a set of verifiable statements, assertions that can be shown to be true with an experimental test in finite time.
+By studying the algebra of such objects, we show that verifiability already provides severe constraints.
+In particular, it requires that a set of physically distinguishable cases is naturally equipped with the mathematical structures (i.e.
+second countable Kolmogorov topologies and $\sigma$-algebras) that form the foundation of manifold theory, differential geometry, measure theory, probability theory and all the major branches of mathematics currently used in physics.
 This gives a clear physical meaning to those mathematical structures and provides a strong justification for their use in science.
-It is our hope and belief that such an approach can be extended to many areas of fundamental physics and will provide a consistent vocabulary across scientific domains.
+It also provides a formal framework to incorporate additional assumptions and constrain the search space for new physical theories.
 \end{abstract}
+
+Clarify our assumptions.
+As we are forced to formalize and make explicit our starting points, we understand under which conditions each physical theory is valid.
+We also have a better idea of when and how they fail, possibly giving new insights that may lead to new physics.
+Put physics back at the center of the discussion.
+By starting from physical ideas we put physics back at the center of the discussion, instead of relegating it to a mere after-the-fact interpretation as it is currently done in most modern scientific theories.
+Many results can be directly understood in terms of the assumptions, fostering a better intuition for the mathematical results.
+Give science sturdier mathematical grounds.
+As each mathematical symbol is given a precise physical meaning, our formal framework maps one-to-one to our physical understanding.
+This means we automatically leave out any unphysical mathematical artifact.
+Foster connections between different fields of knowledge.
+We have found recurring core concepts between different fields of science, math, philosophy and engineering.
+In retrospect, this is not surprising: nature is one and does not care about such divisions, and therefore truly fundamental concepts must span freely across different areas.
 
@@
 
-During the course of the last century a few fields of science and engineering have developed an associated general theory.
-For example, in computer science theory one starts by defining what a computational device is~\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input~\cite{Sipser}.
-In communications one defines what information is~\cite{Shannon} and shows that no encoding can outperform the Huffman coding~\cite{Pierce}.
-In control theory one defines what a system is~\cite{Brogan} and then looks for general strategies for control such as Kalman filtering~\cite{Kalman}.
-These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
-That is, the results apply because of the mere definition of the subject matter.
+Stepping back and looking at physics as a whole, one can't fail to notice that it is essentially composed of different theories and models loosely connected to each other, each with its own starting points and realm of applicability.
+Classical mechanics, electromagnetism, general relativity, quantum mechanics, thermodynamics and so on may share some general ideas but, in the end, they all have their own separate foundation, either in a different set of laws and principles or simply in positing a unique mathematical structure.
+This state of affairs is so entrenched in our field that somehow feels like the best, if not the only, way to go.
+But is it?
+Or this approach actually prevents progress and understanding?
 
-No such general theory currently exists for physics, which is essentially composed of different theories loosely connected to each other, each with their own realm of applicability.
-However, we believe such a general theory can be created and in this work we present our basic strategy towards it.
-As one familiar with the general theories of other subjects would expect, we only need to focus on clarifying exactly what our starting points are and let the logic follow.
-That is, once we formalize exactly what it means to study a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories.
-For example, we will see that any set of objects that are physically distinguishable must have cardinality no greater than continuum.
-We will see that defining what can be experimentally verified about them will always induce a topology that is Kolmogorov and second countable.
-These results are general because they apply independently of what is being studied.
+We can gain perspective comparing to different fields of science and engineering.
+For example, in computer science theory one starts by defining symbolic languages and computational devices~\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input~\cite{Sipser}.
+In communication theory one defines channels and the information they carry~\cite{Shannon} and proves general results such as calculating the maximum rate at which information can be transmitted over a channel given a set amount of noise~\cite{Pierce}.
+In control theory one defines a system in terms of state, inputs and outpust~\cite{Brogan} and then looks for general strategies for control such as Kalman filtering~\cite{Kalman}.
+These theories are \emph{general} in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
+Any topic in each field becomes a specialization of those general mathematical structure.
+In other words, the subject matter has been properly formally defined and the results always apply due to its mere definition.
 
-One should not confuse this approach with the idea of a grand-unified theory or theory of everything.
-In that program, one looks for the theory of fundamental constituents from which all other theories can be derived.
-Our program is altogether different and orthogonal.
-Our subject matter is the scientific models themselves.
-Our aim is to understand what aspects of a scientific theory are already constrained by the mere fact that it is a scientific theory or by the assumptions one builds into them.
-While this general theory may be of some utility to someone working on a grand-unified theory as it may help narrow the search, neither one will lead to the other.
+In physics there is no such thing.
+Classical mechanics, thermodynamics and quantum mechanics are not the specialization of some more general theory.
+But how would that work?
+As in the other general theories, we need to clarify what our starting points are and let the logic follow.
+The first requirement of a physical theory, or rather of a scientific theory in general, is that it is experimentally testable.
+As mathematics is built on the logic of statements, we need to develop the logic of experimental verification.
+This would give the most general mathematical structure that serves as the baseline for everything else.
+The second requirement is having states and processes, which is where physics starts.
+We can then specialize these concepts for different cases.
+Some processes will be deterministic and reversible.
+Some processes will be non-deterministic and some of these will have equilibria.
+By adding different assumptions on states and process we would recover classical Hamiltonian mechanics in one case, quantum mechanics in another and thermodynamics in yet another.
+In all these theories, some basic properties will be common simply because we have states that need to be measured experimentally, and some properties will be different because the type of state or the type of processes are different.
+This is what we mean by a general mathematical theory of experimental science.
 
-In this paper we briefly present the starting point for our general mathematical theory of experimental science.
+Such a general theory would provide a formal framework will force us to clarify our assumptions, clarifying the realm of applicability of each theory.
+It would physics back at the center of our discussion, as physical ideas are the starting point for our derivation and not after-the-fact interpretations.
+It would give science sturdier mathematical grounds, as each mathematical symbol is given a precise meaning and non-physical objects are excluded by construction.
+It would foster connections between different fields of knowledge: nature does not care about such divisions.
+It would provide a framework in which to pose new questions and solve old ones.
+In other words: it would provide a better foundations for physics and for the rest of the sciences.
+Developing a general mathematical theory for experimental science could be comparable to what happened in mathematics during the first half of the last century, when it was reorganized using logic and set theory as its foundation, which had profound repercussions in the field.
+Our preliminary work in both physics~\cite{Carc1} and math~\cite{Carc2} convinced us that such a goal is possible and within reach.
+
+The aim of this work is to lay down the foundations for a general mathematical theory of experimental science.
+We start by defining verifiable statements: assertions whose truth can be tested experimentally in finite time.
+We study the logic of these statements, which is different from the standard Boolean logic because of termination.
+We define experimental domains as a collection of verifiable statement that, if given an indefinite amount of time, we can keep testing forever refining our knowledge.
+From each experimental domain we construct its possibilities: the statements that, if true, would give a prediction to all verifiable statements.
+These represents the different cases we can experimentally distinguish.
+The main results is that \textbf{each set of experimentally distinguishable objects comes equipped with a natural second countable Kolmogorov topology, where each open set corresponds to a verifiable statement, and a natural $\sigma$-algebra, where each Borel set corresponds to a theoretical statement which gives predictions to verifiable ones.} For example, \statement{the mass of the particle is between $0.4$ and $0.6$ MeV} is a verifiable statement precisely because $(0.4, 0.6)$ is an open set, while \statement{the mass of the particle is exactly $0.5$ MeV} is not verifiable because $[0.5,0.5]$ is not an open set.\statement{The standard topology of the real line keeps track of finite precision measurements} From that general result we can derive the following conclusion: \textbf{any set of physically distinguishable cases has most cardinality of continuum}.
+That is, the only thing we can do to distinguish a particular case is run a countable set of tests, whose output can be imagined as a countable sequence of true and false.
+This is equivalent to a real number from 0 to 1 expressed in base 2.
+In a nutshell, \textbf{experimental verification by itself guarantees us the existence of two mathematical structures that are the foundations of most tools used in physics}, such as differential geometry (Riemannian and symplectic), measure theory and probability theory and many other mathematical tools already used in physics and other sciences.
+
+%In this paper we briefly present the starting point for our general mathematical theory of experimental science.
 It is the product of a decade spent in reverse engineering the laws of particle mechanics~\cite{Carc1} and our subsequent work on formalizing the starting points~\cite{Carc2}.
@@
 
-The basic idea is that science deals with verifiable statements: assertions whose truth can be tested experimentally.
-We study the logic of these statements, which is different from the standard Boolean logic.
-We then group them into experimental domains, which represent a set of verifiable statements that can be tested in an indefinite amount of time.
-From these we construct theoretical domains, which represent all possible predictions, and within these we find the possibilities, those predictions that determine the outcome of all possible tests.
-We will find that possibilities have a natural topology induced by the experimental domain and a natural $\sigma$-algebra induced by the theoretical domain.
-We will find that the topology is Kolmogorov and second countable, which limits the cardinality of the set of possibilities of a domain to the one of continuum.
-Note that these structures are indeed pervasive in science, and form the foundation for differential geometry (Riemannian and symplectic), measure theory, probability theory and many other mathematical tools already used in physics and other sciences.
+While this result is remarkable by itself, it provides the basis for future expansions.
+Conceptually, in the same way that topological spaces keep track of what can be verified experimentally, manifolds keep track of objects that can be identified by continuous quantities, differentiable manifolds keep track of objects that allow a density to be defined over them and symplectic manifolds (i.e.
+phase space of classical mechanics) allow the density to be coordinate invariant.
+But by doing all this work formally we are forced to recognize the assumptions that go in those models.
+For example, we have already done preliminary work that uncovered a set of necessary and sufficient conditions such that a set of verifiable statements defines a continuous quantity.
+While, as one would expect, these can only correspond to an idealized case where one pretends a better precision measurement can always be achieved, they also tell us precisely how this idealized case fails which can be interesting for those working on plank scale physics.
+Specifically, ordering must fail which means not only that real number cannot be used to identify configurations in space and time, but that basic mathematical tools such as derivation and integration will break down as well.
+But, whatever other tools we will have, we \emph{know} we will always have a topology and a $\sigma$-algebra.
+
+Another example, note that the set of discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$ has cardinality greater than continuum and therefore cannot represent a set of physically distinguishable objects.
+This means that the square integrable space of functions typically used in quantum mechanics is far too big, and, in fact, it contains unphysical elements, such as states with infinite energy.
+However, the Schwartz space, like any set of continuous functions, has cardinality of continuum and can be given a second countable Kolmogorov topology.\footnote{It also has other properties that may map to physical requirements.
+For example, quantities of the form $\langle \psi|Q^n P^m|\psi\rangle$ will be always finite and the Fourier transform is well defined.} To our understanding, the reason why mathematically Hilbert spaces are useful is because one can take limits and use the vector space norm to converge.
+However, we should be able to use the corresponding $\sigma$-algebra to take limits, like one does in probability and measure theory.
+In other words, we may be able to construct, within this framework, a more mathematically sound and physically motivated foundation for quantum mechanics that includes the objects and only the objects that are physically meaningful.
+In turn, maybe this contribute to a solid mathematical foundation of quantum field theory which, we are told, is still an open problem.
+
+As a last example, we should be able to form a connection to computer science: in the end a computation device is a physical system where the output is verifiable statement that is a function of the inputs, which also must be verifiable statements.
+In this context, all possible inputs form the possibilities of our space.
+Suppose we have a function from input to a Boolean.
+Then we can write the statements \statement{the input is such that the output of the function is true} and \statement{the input is such that the output of the function is false}.
+If both these statements correspond to open sets on the possible inputs, then we construct a test (i.e.
+a program) that will always terminate.
+That is, we can use all the tools in topology and computer science to answer a question of the type: is this problem testable or computable?
+
+We hope that it is clear by now that a general mathematical theory of experimental science is missing, is possible and would be extremely useful.
+Whether you work on condensed matter, quantum thermodynamics, complex systems, particle physics, gravitation or grand unified theories, if what you are doing is science you will have, at some point, experimental verification.
+You will have problems (often the most interesting ones) that arise from questions such as: what is it that I can measure?
+How do I do it?
+Under what assumptions do the quantities in my theory map to those measurements?
+When are they even well defined?
+What can I compute within my theory?
+What is testable?
+Currently, you do not have a general framework to pose those questions.
+The idea that you will need a completely ad-hoc framework for your field only is not true.
+The mere \emph{logic} of experimental verification requires a particular structure.
+And there is a lot more to experimental verification than its logic: the quantity itself has to, at least in a sense, exist over a finite period of time to be measurable, there must be a process such that information about that quantity can be transferred to the measurement device, the system cannot be always assumed to be completely isolated as it would not be physically accessible.
+If you are working on foundational aspects, chances are you may already be thinking about some of these aspects as they relate to your field.
+Our question is: how much of those problems are specific to your field?
+How much can be treated generally, so that on one side each field has less work to do, and on the other side we have a common language across fields?
+What are the assumptions specific to your field and those that are more general?
+A general mathematical theory of experimental science would provide you with a formal framework to precisely pose and possibly answer those questions within your field.
+As the starting points are clarified, the foundations of different fields may be affected, or even partially integrated, and potentially change how they are taught.
+
+The irony is that while the project is extremely ambitious, the biggest obstacles are not, at this point, technical.
+The biggest obstacles are sociological.
+On one side the nature of this work is very interdisciplinary.
+On the other side math and physics are divided in fields and sub-fields that are increasingly narrower.\footnote{We attended a topology conference where there seemed to be a disconnected between what mathematicians and condensed matter physicists meant by topology, to the point that they frequently talked past each other.
+It is also not uncommon to find scientists that work on gravitation that are not familiar with topological spaces, which are the foundation of differential geometry.} This mean that this work does not fit naturally in any community.
+One might think that this work belongs in mathematical physics, because we are working on the fundamental mathematical structures of physics.
+Yet, what most people actually mean by mathematical physics is partial differential equation and differential geometry.
+One might think that this work belongs in foundations of physics, because we are working on making explicit the fundamental assumptions one is taking.
+Yet, what most people actually mean by foundations of physics is interpretations of quantum mechanics, quantum gravity or some other ``quantum thingamajig''.
+This means that this type of work is most often rejected from journals or conferences without review because the editor feels the work is not in scope.
+Or that it is very difficult for journals to find reviewers with the right expertise.
+Or that no funding agency has a category for this type of work.\footnote{Curiously, work on theories that postulate parallel universes or on whether consciousness is what causes particles to decay is somehow well accepted in physics.} Or that arxiv.org categorizes this very paper as artificial intelligence without possibility of appeal.
+Or that, since noone has time and/or expertise to review on the merit, one merely looks at the celebrity of the authors.
+All of this becomes a self-fulfilling prophecy.
+
+A common personal reaction from external observers, in fact, is to characterize this type of work as a new field of study.
+The problem is that normally new fields of study emerge from others as specialization or recombination.
+Therefore there is a time when physics gets discussed as part of natural philosophy, or quantum information is discussed as part of quantum mechanics or statistical mechanics.
+If sufficient critical mass is established, new fields, journal and funding categories are created.
+In our case, we are taking elements across too many different disciplines and putting them together to create something broader for each discipline to recognize itself in it.
+So we are most often told that our work \emph{is} interesting and worthwhile and it \emph{should} be done, but it should be done \emph{somewhere else} by \emph{someone else}.
+It is a strange form of Not In My Back Yard (NIMBY) that hits many who try to do interdisciplinary work and leaves us with no temporary home.
+
+To obviate to these problems, borrowing practices from the open source software community, we are developing our main body of work in draft open access book form~\cite{Carc3}, so that we can address the topic in its entirety and our friends in philosophy, mathematics and physics can all check their parts.
+In fact, this very article is derived from that work and one should refer to it to find more mathematical and philosophical details.
+This type of peer review process is what is needed to get the job done, but it is impossible to do within traditional academic settings.
+It is only possible because we have found experts with very different backgrounds who value the work and understand its potential.
+For that we are lucky.
 
@@
 
-	\bibitem{Carc2} C. A. Aidala, G. Carcassi, M. J. Greenfield: Topology and experimental distinguishability. arXiv:1708.05492, 2017.
+	\bibitem{Carc2} C. A. Aidala, G. Carcassi, M. J. Greenfield: Topology and experimental distinguishability. Accepted by Topology Proceedings on 11 Dec 2018, arxiv:1708.05492.
 

### 2019-02-08 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Reviewed introduction and conclusions

--- 2018-06-22 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
@@
+\newif\ifarchive
+\archivetrue
+%\archivefalse
+
+\ifarchive
 	\documentclass[letterpaper]{article}
+\else
+	\documentclass[smallextended]{svjour3}
+\fi
 
 %% Language and font encodings
-\usepackage[english]{babel}
-\usepackage[utf8x]{inputenc}
-\usepackage[T1]{fontenc}
+%\usepackage[english]{babel}
+%\usepackage[utf8x]{inputenc}
+%\usepackage[T1]{fontenc}
 
 %% Sets page size and margins
-\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}
+%\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}
 
-% Temporary to remove "submitted to"
-\makeatletter
-\def\ps@pprintTitle{%
-	\let\@oddhead\@empty
-	\let\@evenhead\@empty
-	\def\@oddfoot{}%
-	\let\@evenfoot\@oddfoot}
-\makeatother
+%\usepackage{lineno,hyperref}
+\usepackage{tikz}
 
-\usepackage{lineno,hyperref}
-
+\ifarchive
+% Theorem definitions using amsthm
 \usepackage{amsmath, amsthm, amsfonts}
-\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}
 
@@
 \begin{abstract}
-	In this article we lay the groundwork for a general mathematical theory of experimental science.
-The starting point will be the notion of verifiable statements, those assertions that can be shown to be true with an experimental test.
-We study the algebra of such objects and show how it is closed only under finite conjunction and countable disjunction.
-With simple constructions, we show that the set of possible cases distinguishable by verifiable statements is equipped with a natural Kolmogorov and second countable topology and a natural $\sigma$-algebra.
+	We lay the groundwork for a formal framework that studies scientific theories and can serve as a unified foundation for the different theories within physics.
+We define a scientific theory as a set of verifiable statements, assertions that can be shown to be true with an experimental test in finite time.
+By studying the algebra of such objects, we show that verifiability already provides severe constraints.
+In particular, it requires that a set of physically distinguishable cases is naturally equipped with the mathematical structures (i.e.
+second-countable Kolmogorov topologies and $\sigma$-algebras) that form the foundation of manifold theory, differential geometry, measure theory, probability theory and all the major branches of mathematics currently used in physics.
 This gives a clear physical meaning to those mathematical structures and provides a strong justification for their use in science.
-It is our hope and belief that such an approach can be extended to many areas of fundamental physics and will provide a consistent vocabulary across scientific domains.
+Most importantly it provides a formal framework to incorporate additional assumptions and constrain the search space for new physical theories.
 \end{abstract}
-
-% We form experimental domains, which represent the list of all possible testable answers to a particular scientific question.
-From these we construct theoretical domains, which represent all possible predictions (not necessarily verifiable themselves).
-We then define the possibilities of an experimental domain as those predictions that, if true, predict the truthfulness of all verifiable statements in the domain.
-We then show that experimental domains provide a natural Kolmogorov and second countable topology over their possibilities, while theoretical domains provide a natural $\sigma$-algebra.
 
@@
 
-During the course of the last century a few fields of science and engineering have developed an associated general theory.
-For example, in computer science theory one starts by defining what a computational device is~\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input~\cite{Sipser}.
-In communications one defines what information is~\cite{Shannon} and shows that no encoding can outperform the Huffman coding~\cite{Pierce}.
-In control theory one defines what a system is~\cite{Brogan} and then looks for general strategies for control such as Kalman filtering~\cite{Kalman}.
-These theories are general in the sense that their results respectively apply to all control systems, all communication systems and all computational devices.
-That is, the results apply because of the mere definition of the subject matter.
+When considering physics as a discipline, one cannot help but notice that it is essentially composed of different theories and models loosely connected to each other, each with its own starting points and realm of applicability.
+Classical mechanics, electromagnetism, general relativity, quantum mechanics, thermodynamics and so on may share some general ideas but, in the end, they all have their own separate foundation, either in a different set of laws and principles or simply in positing a unique mathematical structure.
+This state of affairs is so entrenched in our field it somehow feels like the proper, if not the only, approach.
+But is it?
+Or does this approach actually hinder progress and understanding?
 
-No such general theory currently exists for physics, which is essentially composed of different theories loosely connected to each other, each with their own realm of applicability.
-However, we believe such a general theory can be created and in this work we present our basic strategy towards it.
-As one familiar with the general theories of other subjects would expect, we only need to focus on clarifying exactly what our starting points are and let the logic follow.
-That is, once we formalize exactly what it means to study a system through experimentation, we can reach interesting results that necessarily apply to all scientific theories.
-For example, we will see that any set of objects that are physically distinguishable must have cardinality no greater than continuum.
-We will see that defining what can be experimentally verified about them will always induce a topology that is Kolmogorov and second countable.
-These results are general because they apply independently of what is being studied.
+We can gain perspective by comparing to different fields of science and engineering.
+For example, in computer science theory one starts by defining symbolic languages and computational devices~\cite{Turing} and then shows that no algorithm exists that can correctly decide whether an arbitrary program terminates given an arbitrary input~\cite{Sipser}.
+In communication theory one defines channels and the information they carry~\cite{Shannon} and proves general results such as calculating the maximum rate at which information can be transmitted over a channel given a set amount of noise~\cite{Pierce}.
+In control theory one defines a system in terms of state, inputs and outputs~\cite{Brogan} and then looks for general strategies for control such as Kalman filtering~\cite{Kalman}.
+These theories, with their respective mathematical structures and results, are \emph{general} in the sense that they apply to all control systems, all communication systems and all computational devices.
+More specific topics in each field are constructed by further constraining those general mathematical structures.
+In other words, the subject matter has been properly defined formally and therefore the results follow simply from the mere definition of the problem.
 
-One should not confuse this approach with the idea of a grand-unified theory or theory of everything.
-In that program, one looks for the theory of fundamental constituents from which all other theories can be derived.
-Our program is altogether different and orthogonal.
-Our subject matter is the scientific models themselves.
-Our aim is to understand what aspects of a scientific theory are already constrained by the mere fact that it is a scientific theory or by the assumptions one builds into them.
-While this general theory may be of some utility to someone working on a grand-unified theory as it may help narrow the search, neither one will lead to the other.
+In physics there is no equivalent.
+Classical mechanics, thermodynamics and quantum mechanics are not specific topics within some more general formal theory.
+But how would that work?
+As in the other general theories, we need to clarify what our starting points are and let the logic follow.
+The first requirement of a physical theory, or rather of any scientific theory in general, is that it is experimentally testable.
+So we would start by characterizing the properties and limitations of experimental verification.
+The second requirement, which is where physics starts, is identifying states and processes.
+After those are characterized in general, we can then specialize them for different cases.
+Some processes will be deterministic and reversible.
+Some processes will be non-deterministic and some will have equilibria.
+Some states can be broken into parts and some cannot.
+By adding different assumptions on states and processes we would recover classical Hamiltonian mechanics in one case, quantum mechanics in another and thermodynamics in yet another.
+In all these theories, some basic properties will be common simply because states need to be identified experimentally, and some properties will be different because the type of state or the type of process is different.
+This is what we mean by a general mathematical theory of experimental science.
 
-In this paper we briefly present the starting point for our general mathematical theory of experimental science.
+%[rectangle, draw, fill=black!10, inner sep=5pt, text width=4cm, text badly centered, minimum height=1.2cm]
+%every node/.style={draw, align = center},
+
+\begin{figure}\label{twoQuestions}
+	\centering
+	\begin{tikzpicture}
+	[child anchor = north, every node/.style={draw, align = center, inner sep=5pt},
+	edge from parent/.style={draw, ->, -latex, stealth-},
+	level 1/.style={level distance=1cm, sibling distance=6cm},
+	level 2/.style={level distance=1.3cm, sibling distance=3cm}, sloped]
+	\node {Experimental verifiability} [text width=4cm]
+	child {node {States and processes} [text width={}]
+		child {
+			node {Classical \\ mechanics}
+		}
+		child {
+			node {Quantum \\ mechanics}
+		}
+		child {
+			node {Thermodynamics}
+		}
+		child {
+			node {...}
+		}
+	};
+	\end{tikzpicture}
+	\caption{Overall structure for a general mathematical theory of experimental science.
+The starting point is experimental verification.
+States and processes are particular types of experimentally testable objects.
+Different theories in physics describe particular states under particular processes.}
+\end{figure}
+
+Such a general theory would force us to clarify our assumptions, thus the realm of applicability of each theory.
+It would always put physics at the center of our discussion, as physical ideas are the starting point for our formal framework and not after-the-fact interpretations.
+It would give science sturdier mathematical grounds, as each mathematical symbol is given a precise meaning and non-physical objects are excluded by construction.
+It would foster connections between different fields of knowledge: nature does not care about such divisions.
+It would provide a framework in which to pose new questions and solve old ones.
+In other words: it would provide a better foundation for physics and for the rest of the sciences.
+Developing a general mathematical theory for experimental science could be compared to what happened in mathematics during the first half of the last century, when it was reorganized using logic and set theory as its foundation, which had profound repercussions in the field.
+Our preliminary work in both physics~\cite{Carc1} and math~\cite{Carc2} convinced us that such a goal is possible and within reach.
+
+The aim of the present work is to lay down the beginning of a general mathematical theory of experimental science, limited to the part concerning experimental verification.
+We start by defining verifiable statements: assertions whose truth can be verified experimentally in finite time.
+We study the logic of these statements, which is different from the standard Boolean logic because of finite time termination.
+We define experimental domains as a collection of verifiable statements that, if given an indefinite amount of time, we can keep testing, forever refining our knowledge.
+From each experimental domain we construct its possibilities: the different cases we can experimentally distinguish.
+The main result is that \textbf{each set of experimentally distinguishable possibilities comes equipped with a natural second-countable Kolmogorov topology, where each open set corresponds to a verifiable statement, and a natural $\sigma$-algebra, where each Borel set corresponds to a theoretical statement which gives predictions for verifiable ones.} For example, \statement{the mass of the particle is more than $0.4$ and less than $0.6$ MeV} is a verifiable statement precisely because $(0.4, 0.6)$ is an open set, while \statement{the mass of the particle is exactly $0.5$ MeV} is not verifiable because $[0.5,0.5]$ is not an open set.\footnote{The standard topology of the real numbers keeps track of finite precision measurements.} From that general result we can derive the following conclusion: \textbf{any set of physically distinguishable cases has at most cardinality of the continuum}.
+That is, the only thing we can do to distinguish a particular case is run the tests for a countable set of statements, the output of which can be imagined as a countable sequence of true and false.
+This is equivalent to the binary representation of a real number.
+In a nutshell, \textbf{experimental verification by itself guarantees us the existence of two mathematical structures that are the foundations of most tools used in physics}, such as differential geometry (Riemannian and symplectic), measure theory, probability theory and many others.
+
+%In this paper we briefly present the starting point for our general mathematical theory of experimental science.
 It is the product of a decade spent in reverse engineering the laws of particle mechanics~\cite{Carc1} and our subsequent work on formalizing the starting points~\cite{Carc2}.
@@
 
-The basic idea is that science deals with verifiable statements: assertions whose truth can be tested experimentally.
-We study the logic of these statements, which is different from the standard Boolean logic.
-We then group them into experimental domains, which represent a set of verifiable statements that can be tested in an indefinite amount of time.
-From these we construct theoretical domains, which represent all possible predictions, and within these we find the possibilities, those predictions that determine the outcome of all possible tests.
-We will find that possibilities have a natural topology induced by the experimental domain and a natural $\sigma$-algebra induced by the theoretical domain.
-We will find that the topology is Kolmogorov and second countable, which limits the cardinality of the set of possibilities of a domain to the one of continuum.
-Note that these structures are indeed pervasive in science, and form the foundation for differential geometry (Riemannian and symplectic), measure theory, probability theory and many other mathematical tools already used in physics and other sciences.
+While this result is remarkable by itself, the most interesting aspect is that it can be used as a cornerstone for a more general theory.
+Conceptually, in the same way that topological spaces keep track of what can be verified experimentally, manifolds keep track of objects that can be identified by continuous quantities, differentiable manifolds keep track of objects that allow a density to be defined over them and symplectic manifolds (i.e.
+phase space of classical mechanics) allow the density to be coordinate invariant (i.e.
+observer independent).
+By doing all this work formally we will be forced to recognize the assumptions that go into those structures, and understand if and how they fail.
+For example, we are currently in the process of identifying the necessary and sufficient conditions such that a set of verifiable statements defines a continuous quantity.
+While, as one would expect, these can only correspond to an idealized case where one pretends a better precision measurement can always be achieved, they also tell us precisely how this idealized case fails which can be relevant to work on Planck scale physics.
+But whatever other tools we will have in that case, we \emph{know} we will always have at least a topology and a $\sigma$-algebra.
+
+As another example, note that the set of discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$ has cardinality greater than the continuum and therefore cannot represent a set of physically distinguishable objects.
+This means that the square integrable space of functions typically used in quantum mechanics is far too big, and, in fact, it contains unphysical elements, such as states with infinite energy.
+However, the Schwartz space, like the set of continuous functions, has cardinality of the continuum and can be given a second-countable Kolmogorov topology.\footnote{It also has other properties that map to physical requirements.
+For example, quantities of the form $\langle \psi|Q^n P^m|\psi\rangle$ are always finite and the Fourier transform is always well defined.} To our understanding, the reason why Hilbert spaces are mathematically useful is because one can take limits and use the vector space norm to converge.
+However, we may instead be able to use the corresponding $\sigma$-algebra to take limits, like one does in probability and measure theory.
+In other words, we may be able to construct, within this framework, a more mathematically sound and physically motivated foundation for quantum mechanics that includes the objects and only the objects that are physically meaningful.
+
+As a last example, we also believe we can form a connection to computer science.
+We can characterize the output of a computational device as a set of verifiable statements (e.g.
+\statement{the third bit of output is 1}) that are a function of the input, which can also be characterized as verifiable statements (e.g.
+\statement{the second bit of the input is 0}).
+In this context, all possible inputs form the possibilities of our space, as they define all possible outputs of the computation.
+Now suppose we have a function from the input to a Boolean value.
+Then we can write the statements \statement{the input is such that the output of the function is true} and \statement{the input is such that the output of the function is false}.
+If both these statements correspond to open sets on the possible inputs then they are verifiable statements and we construct a test (i.e.
+a program) that will always terminate.
+That is, we can combine tools from topology and computer science to answer a question of the type: is this scientific problem testable or computable?
+In a similar way, connections to information theory can also be established: a deterministic and reversible process, in fact, is one for which description of the input is equivalent to the description of the output, which means information entropy is conserved.
+These links illustrate the potential as a theory of experimental science, and not just physics.
+
+We hope that it is clear by now that a general mathematical theory of experimental science is missing, is possible and would be extremely useful.
+Whether you work on condensed matter, quantum thermodynamics, complex systems, particle physics, gravitation or grand unified theories, if what you are doing is science you will have, at some point, experimental verification.
+You will have problems (often the most interesting ones) that arise from questions such as: what is it that I can measure?
+How do I do it?
+Under what assumptions do the quantities in my theory map to those measurements?
+When are they even well defined?
+What can I compute within my theory?
+What is testable?
+Currently, you do not have a general framework to pose those questions.
+And it turns out you do not need a completely ad-hoc one for your specific field.
+The mere \emph{logic} of experimental verification requires a particular structure.
+And there is a lot more to experimental verification than its logic: the quantity itself has to, at least in a sense, exist over a finite period of time to be measurable, there must exist a process to transfer that information to a measurement device, the system cannot be assumed to be always completely isolated as it would not be physically accessible.
+If you are working on foundational aspects, chances are you may already be thinking about some of these issues as they relate to your field.
+But what parts of those problems are specific to your field?
+How much can be treated generally, so that not only each field has less work to do but also we have a common language across fields?
+What are the assumptions specific to your field and those that are more general?
+A general mathematical theory of experimental science would provide you with a formal framework to precisely pose and possibly answer those questions.
+As the starting points are clarified, the foundations of different fields may be affected, or even partially integrated, and potentially change how they are taught as well.
+
+The irony is that while the project is extremely ambitious, the biggest obstacles are not, at this point, technical.
+The biggest obstacles are sociological.
+On one side the nature of this work is extremely interdisciplinary.
+On the other side math and physics are divided into fields and sub-fields that are increasingly narrow.\footnote{We attended a topology conference where there seemed to be a disconnect between what mathematicians and condensed matter physicists meant by topology, to the point that they frequently talked past each other.} This means that this work does not fit naturally within any community while requiring technical expertise from many.
+To obviate these problems, borrowing practices from the open source software community, we are developing our main body of work in the open~\cite{Carc3}, so that we can address the topic in its entirety, and our colleagues in philosophy, mathematics and physics can all check their respective parts.
+This article is the product of that process and focuses on current results of greater interest to physicists, omitting more mathematical and philosophical details that can be found in the broader work.
 
@@
 
-The first thing we need to do is to characterize what science is and what its scope is. We do so by introducing the following:
+As the starting point of our general theory, we introduce the following guiding principle:
 
@@
 
-This means that science restricts itself to the study of assertions that have a well defined truth value which can be verified experimentally.
+This means that science concerns itself only to the study of assertions that have a well defined truth value which can be verified experimentally.
 The issue at hand is to formally capture this informal intuition.
-We start with the following common definition.
+We begin with the following common definition.
 
@@
 For example, \statement{this animal is a dog} and \statement{questo animale \`e un cane} represent the same fact expressed in different languages.
-As science is universal, it should not matter the language, units or reference system used to make an assertion.
-While we will still use standard mathematical logic for the formal system, we introduce a variation of algebraic logic to represent our ``informal'' assertions.
+As science is universal, it should not matter the language, units or reference system used to make an assertion.\footnote{In the same vein, a statement is not necessarily one sentence as it has to include all the information needed to understand what exactly we are taking about.
+These are the type of philosophical issues that are discussed at greater length in~\cite{Carc3}.} While we will still use standard mathematics for the formal system, we introduce a variation of algebraic logic to represent our ``informal'' assertions.
 We start by defining our truth bearer as:
@@
 
-\section{Brief discussion}
+\section{Brief discussion and examples}
 
@@
 Every theorem, every proof on those structures can now be given a direct physical meaning as well.
-And given the foundational nature of topologies and $\sigma$-algebras, we hope to extend this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.
+And given the foundational nature of topologies and $\sigma$-algebras, we can imagine extending this framework to measure theory, differential geometry, symplectic geometry, Riemannian geometry, probability theory and so on.
 
-To give an example, consider an experimental domain where the basis is composed of statements like \statement{this quantity is between $a$ and $b$} where $a$ and $b$ are different values that can be arbitrarily close.
-For example, \statement{the distance between the earth and the moon is between 384 and 385 thousand Km}.
+To give an example of how it works, consider an experimental domain where the basis is composed of statements like \statement{this quantity is more than $x$ but less than $y$} where $x$ and $y$ are different values that can be arbitrarily close.
+For example, \statement{the distance between the earth and the moon is more than 384 but less than 385 thousand Km}.
 Each verifiable statement of the basis corresponds to an open interval of the real line, therefore we find a correspondence between arbitrary precision measurements and the standard topology on the reals, as this is the one generated by open intervals.
-Note, in fact, that this topology is second-countable and Kolmogorov.
+Note, in fact, that this topology is second-countable and at least Kolmogorov.
 
-The possibilities (i.e.
+The possibilities (e.g.
 \statement{the mass of the photon is precisely zero}) are not themselves verifiable since infinite precision measurements of a continuous quantity are not possible.
 Yet, their negation (i.e.
-\statement{the mass of the photon isn't precisely zero}) could be verified in practice.
+\statement{the mass of the photon is not precisely zero}) could be verified in practice.
 This is expressed mathematically by the fact that singletons are closed sets.
-The theoretical domain, instead, corresponds to the standard Borel algebra.
+The theoretical domain, instead, corresponds to the standard Borel algebra, which includes closed and half-open sets but not all possible sets of integers, many of which cannot be characterized by a well formed formula.
 
-Similarly, one can imagine verifiable statements for relationships (i.e.
-\statement{when the temperature of the mercury column is between 24 and 25 C, its height is between 24 and 25 mm}) and for statistical variables (i.e.
-\statement{if this coin is tossed enough times, the ratio of heads will be between 45\% and 55\%}).
+Similarly, one can imagine verifiable statements for relationships (e.g.
+\statement{when the temperature of the mercury column is between 24 and 25 C, its height is between 24 and 25 mm}) and for statistical variables (e.g.
+\statement{if this coin is tossed enough times, the fraction of heads will be between 45\% and 55\%}).
 Ultimately the general theory will need to show precisely what mathematical structures map to experimental domains formed by statements of these types and under what assumptions.
+But the idea is that, once you have specified the verifiable statements and their logical relationships, you have already specified the experimental domain and nothing else needs to be added.
+In other words: \textbf{the points of the space and their mathematical structures are fully specified by what can be measured within the theory}.
 
@@
 
-The general theory presented in this paper can be briefly summarized as follows.
-Science deals with statements that can be shown to be true experimentally.
-A scientific domain (i.e.
-an experimental domain) will be defined by a set of verifiable statements.
-From these we construct the set of all possible predictions (i.e.
-the theoretical domain), and concentrate on those that fully predict all our measurements, the possibilities.
-By the very nature and properties of these objects, a topology and a $\sigma$-algebra is induced on the possibilities that retains the connection to which statements are verifiable and which statements constitute a valid prediction.
+Science is based on experimental verification.
+Experimental verification has its own logic.
+This logic imposes a mathematical structure which we characterized in terms of experimental domains (collections of verifiable statements), theoretical domains (collections of predictions for verifiable statements) and possibilities (the cases that can be distinguished experimentally).
+This mathematical structure leads naturally to topological spaces and $\sigma$-algebras, the foundation of many of the tools used in physics and science in general.
+Not only does this result clarify why those tools are so successful in science, it provides a direct physical meaning to those structures and a solid foundation upon which to create a general mathematical theory of experimental science.
 
-The framework will be further expanded by studying how additional assumptions constrain the verifiable statements.
-Assuming a quantity can be measured with arbitrary precision will lead to the real numbers with the standard topology.
-Similarly, other assumptions will lead to other mathematical structures.
-This expansion is the aim of future work.
+In this light, what is most remarkable about this work is not its results, but that \emph{it can be done in the first place}.
+That there is a way to construct physical theories that forces us to spell out our physical assumptions, and that clarifies in a rigorous way what result is a consequence of what assumption.
+We can therefore analyze and compare new starting points with the discipline and thoroughness of modern mathematics instead of just rummaging through the bag of mathematical tools.
+But to do that, we have to create a space in the scientific community that is conducive to this kind of broad interdisciplinary work.
+We need generalists that can recognize how a detail in measure theory relates to information entropy or Hamiltonian mechanics.
+We need to take the mathematical structures created by mathematicians for their needs, break them apart and recombine them to find mathematical structures most suited and meaningful for science.
+Scientific knowledge has expanded quite considerably in the last century, maybe it is time for a moment of synthesis and consolidation.
 
-One benefit of the general theory is to bring to the forefront our hidden assumptions.
-If real numbers are associated with arbitrary precision, what happens in the case of limited precision, when we cannot assume precision can always be increased indefinitely?
-Would we still have real numbers?
-Would we be justified to use smooth manifolds or would we need an entirely different topological structure?
-
-The other benefit is that no interpretation is required.
-All mathematical definitions are capturing well defined physical concepts.
-Therefore every theorem, every step in every proof will also have a well defined physical meaning.
-
-We feel that this type of work is of significance for the scientific community in general and for the physics community in particular if we are ever to truly understand what scientific theories are meant to be describing.
+We believe the development of a general mathematical theory of experimental science would be an important accomplishment for the scientific community and that it would be beneficial in ways we cannot yet even imagine.
+All we know is that a clearer understanding of our starting points cannot but help to provide insights into the current theories and suggest strategies to address long standing problems.
 

### 2019-03-12 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
As submitted to Rev X

@@
 
-\ifarchive
-	\documentclass[letterpaper]{article}
-\else
-	\documentclass[smallextended]{svjour3}
-\fi
+\documentclass[%
+reprint,
+%superscriptaddress,
+%groupedaddress,
+%unsortedaddress,
+%runinaddress,
+%frontmatterverbose, 
+%preprint,
+%preprintnumbers,
+%nofootinbib,
+%nobibnotes,
+%bibnotes,
+amsmath,amssymb,
+aps,
+%pra,
+prx,
+%rmp,
+%prstab,
+%prstper,
+%floatfix,
+]{revtex4-1}
 
@@
 
-\ifarchive
 % Theorem definitions using amsthm
@@
 \newtheorem*{rem}{Remark}
-
-\else
-
-\usepackage{amsmath, amsfonts}
-
-\numberwithin{equation}{section}
-\spnewtheorem{assump}{Assumption}{\bf}{\it}
-\renewcommand{\theassump}{\Roman{assump}}
-\spnewtheorem{prop}[equation]{Proposition}{\bf}{\rm}
-\spnewtheorem{thrm}[equation]{Theorem}{\bf}{\it}
-\spnewtheorem{defn}[equation]{Definition}{\bf}{\it}
-\spnewtheorem{axiom}[equation]{Axiom}{\bf}{\it}
-\spnewtheorem*{principle*}{Principle of scientific objectivity}{\bf}{\it}
-
-\newenvironment{rationale}{\emph{Rationale}.}{\hfill\(\qed\)}
-\newenvironment{justification}{\emph{Justification}.}{\hfill\(\qed\)}
-\renewenvironment{proof}{\emph{Proof}.}{\hfill\(\qed\)}
-\fi
 
@@
 
-\ifarchive
-\author{Gabriele Carcassi, Christine A. Aidala \\ University of Michigan}
+\author{Gabriele Carcassi}
+\email{carcassi@umich.edu}
+\author{Christine A. Aidala}
+\affiliation{
+	Physics Department \\
+	University of Michigan \\
+	Ann Arbor, MI 48109 \\
+}
+
 \date{\today}
-\else
-\author{Gabriele Carcassi \and Christine A. Aidala}
-\institute{G. Carcassi, C. A. Aidala \at
-	Physics Department, University of Michigan, Ann Arbor, MI 48109 \\
-	\email{carcassi@umich.edu}
-}
-\date{Received: date / Accepted: date}
-\fi
-
-\maketitle
 
@@
 \end{abstract}
+
+\maketitle
 
@@
 	edge from parent/.style={draw, ->, -latex, stealth-},
-	level 1/.style={level distance=1cm, sibling distance=6cm},
-	level 2/.style={level distance=1.3cm, sibling distance=3cm}, sloped]
-	\node {Experimental verifiability} [text width=4cm]
-	child {node {States and processes} [text width={}]
+	level 1/.style={level distance=1cm},
+	level 2/.style={level distance=1.3cm, sibling distance=2.4cm}, sloped]
+	\node {Experimental verifiability}
+	child {node {States and processes}
 		child {
@@
 		child {
-			node {Thermodynamics}
+			node {Thermo- \\dynamics}
 		}
@@
 However, the Schwartz space, like the set of continuous functions, has cardinality of the continuum and can be given a second-countable Kolmogorov topology.\footnote{It also has other properties that map to physical requirements.
-For example, quantities of the form $\langle \psi|Q^n P^m|\psi\rangle$ are always finite and the Fourier transform is always well defined.} To our understanding, the reason why Hilbert spaces are mathematically useful is because one can take limits and use the vector space norm to converge.
+For example, quantities of the form \unexpanded{$\langle \psi|Q^n P^m|\psi\rangle$} are always finite and the Fourier transform is always well defined.} To our understanding, the reason why Hilbert spaces are mathematically useful is because one can take limits and use the vector space norm to converge.
 However, we may instead be able to use the corresponding $\sigma$-algebra to take limits, like one does in probability and measure theory.
@@
 For example, \statement{this animal is a dog} and \statement{questo animale \`e un cane} represent the same fact expressed in different languages.
-As science is universal, it should not matter the language, units or reference system used to make an assertion.\footnote{In the same vein, a statement is not necessarily one sentence as it has to include all the information needed to understand what exactly we are taking about.
+As science is universal, it should not matter the language, units or reference system used to make an assertion.\footnote{In the same vein, a statement is not necessarily one sentence as it has to include all the information needed to understand what exactly we are talking about.
 These are the type of philosophical issues that are discussed at greater length in~\cite{Carc3}.} While we will still use standard mathematics for the formal system, we introduce a variation of algebraic logic to represent our ``informal'' assertions.
@@
 
-We could also define decidable statements as those for which falsehood can also be tested experimentally. This table compares the different algebras.
+We could also define decidable statements as those for which falsehood can also be tested experimentally. Table \ref{tab:algebras} compares the different algebras.
 
-\begin{table}[h]
+\begin{table*}
 	\centering
@@
 	\end{tabular}
-	\caption{Comparing algebras of statements.}
-\end{table}
+	\caption{Comparing algebras of statements.}\label{tab:algebras}
+\end{table*}
 
@@
 This means that we are never going to be able to experimentally distinguish between elements of greater cardinality.
-The set of topologically discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$, for example, has greater cardinality and therefore it will never be associated with any physically distinguishable concept.\footnote{However, discontinuous function with up to countably many discontinuities can be seen as the limit of a sequence of continuous functions and therefore may be used as a convenient approximation of a physical concept.} All the issues with large cardinals are not something science will ever be interested in.
+The set of topologically discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$, for example, has greater cardinality and therefore it will never be associated with any physically distinguishable concept.\footnote{However, a discontinuous function with up to countably many discontinuities can be seen as the limit of a sequence of continuous functions and therefore may be used as a convenient approximation of a physical concept.} All the issues with large cardinals are not something science will ever be interested in.
 It does not matter what system we are describing, what experimental techniques we are using or how clever we are.
@@
 \section{Topologies and sigma-algebras}
+
+\begin{table*}
+	\centering
+	\begin{tabular}{p{0.075\textwidth} p{0.275\textwidth} p{0.2\textwidth} p{0.3\textwidth}}
+		& Statement relationship & & Set relationship  \\ 
+		\hline 
+		$\stmt_1 \AND \stmt_2$ & (Conjunction) & $U(\stmt_1) \cap U(\stmt_2)$ & (Intersection) \\ 
+		$\stmt_1 \OR \stmt_2$ & (Disjunction) & $U(\stmt_1) \cup U(\stmt_2)$ & (Union) \\ 
+		$\NOT \stmt$ & (Negation) & $U(\stmt)^C$ & (Complement) \\ 
+		$\stmt_1 \equiv \stmt_2$ & (Equivalence) & $U(\stmt_1) = U(\stmt_2)$ & (Equality) \\ 
+		$\stmt_1 \narrower \stmt_2$ & (Narrower than) & $U(\stmt_1) \subseteq U(\stmt_2)$ & (Subset) \\ 
+		$\stmt_1 \broader \stmt_2$ & (Broader than) & $U(\stmt_1) \supseteq U(\stmt_2)$ & (Superset) \\ 
+		$\stmt_1 \comp \stmt_2$ & (Compatibility) & $U(\stmt_1) \cap U(\stmt_2) \neq \emptyset$ & (Intersection not empty)
+	\end{tabular} 
+	\caption{Correspondence between statement operators and set operators.}\label{tab:statement_set}
+\end{table*}
 
@@
 
-Since every statement is a set of possibilities, we can re-express the statement relationships in terms of set relationships according to this table:
-
-\begin{table}[h]
-	\centering
-	\begin{tabular}{p{0.075\textwidth} p{0.275\textwidth} p{0.2\textwidth} p{0.3\textwidth}}
-		& Statement relationship & & Set relationship  \\ 
-		\hline 
-		$\stmt_1 \AND \stmt_2$ & (Conjunction) & $U(\stmt_1) \cap U(\stmt_2)$ & (Intersection) \\ 
-		$\stmt_1 \OR \stmt_2$ & (Disjunction) & $U(\stmt_1) \cup U(\stmt_2)$ & (Union) \\ 
-		$\NOT \stmt$ & (Negation) & $U(\stmt)^C$ & (Complement) \\ 
-		$\stmt_1 \equiv \stmt_2$ & (Equivalence) & $U(\stmt_1) = U(\stmt_2)$ & (Equality) \\ 
-		$\stmt_1 \narrower \stmt_2$ & (Narrower than) & $U(\stmt_1) \subseteq U(\stmt_2)$ & (Subset) \\ 
-		$\stmt_1 \broader \stmt_2$ & (Broader than) & $U(\stmt_1) \supseteq U(\stmt_2)$ & (Superset) \\ 
-		$\stmt_1 \comp \stmt_2$ & (Compatibility) & $U(\stmt_1) \cap U(\stmt_2) \neq \emptyset$ & (Intersection not empty)
-	\end{tabular} 
-	\caption{Correspondence between statement operators and set operators.}
-\end{table}
+Since every statement is a set of possibilities, we can re-express the statement relationships in terms of set relationships according to Table \ref{tab:statement_set}.
 
@@
 We thank Mark J.
-Greenfield for the thorough review of the mathematical framework.
-We thank Josh Hunt for comments on the philosophical underpinning of this work.
-We would also like to acknowledge the numerous individuals who commented on the drafts of this paper and our previous works leading up to it.
+Greenfield for the thorough scrutiny of the mathematical framework and Mathew Timm for clarifying what can and cannot be defined within a formal mathematical framework.
+We thank Josh Hunt for the detailed review of the philosophical underpinning of this work and Gordon Belot, Laura Ruetsche and David J.
+Baker for philosophical discussions related to this work.
+We would also like to acknowledge John Mayer, Kai Sun, Jens Zorn and others for their interest and support for this project.
 
@@
 
-\begin{thebibliography}{1}
-
-	\bibitem{Turing}
-	A. M. Turing: On Computable Numbers, with an Application to the Entscheidungsproblem". Proceedings of the London Mathematical Society 2, 42: 230–265, doi:10.1112/plms/s2-42.1.230, 1936.
-
-	\bibitem{Sipser} M. Sipser: An Introduction to the Theory of Computation, 3rd edition. Cengage Learning, Boston, USA, 2013.
-
-	\bibitem{Shannon} C. E. Shannon: A Mathematical Theory of Communication. The Bell System Technical Journal,
-	Vol. 27, pp. 379–423, 623–656, July, October, 1948.
-
-	\bibitem{Pierce} J. R. Pierce: An Introduction To Information Theory: Symbols, Signals and Noise. Dover Books on Mathematics, Second Edition, 1980.
-
-	\bibitem{Brogan} W. L. Brogan: Modern Control Theory, 3rd edition. Pearson, 1990.
-
-	\bibitem{Kalman} R. E. Kalman: A new approach to linear filtering and prediction problems. Journal of Basic Engineering. 82 (1): 35–45, doi:10.1115/1.3662552, 1960.
-
-	\bibitem{Carc1} G.
-Carcassi, C.
-A.
-Aidala, D.
-J.
-Baker and L.
-Bieri: From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics.
-Journal of Physics Communications, 2, 4, 045026, 2018.
-
-	\bibitem{Carc2} C. A. Aidala, G. Carcassi, M. J. Greenfield: Topology and experimental distinguishability. Accepted by Topology Proceedings on 11 Dec 2018, arxiv:1708.05492.
-
-	\bibitem{Carc3} G. Carcassi, C. A. Aidala: Assumptions of Physics (in preparation). http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.
-\end{thebibliography}
-
 \end{document}

### 2019-03-14 Articles/2019-ForQuantumOptics/article.tex
First draft for first two sections

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{dutchcal}
\usepackage{braket}
\usepackage{enumitem}

\usepackage{tikz}
\usepackage{forest}
\usetikzlibrary{trees}
\usetikzlibrary{calc}
\usepackage{calculator}
\usepackage{standalone}

\usepackage{braket}

\begin{document}

\title{TODO}

\maketitle

\begin{abstract}
	TODO
\end{abstract}

\tableofcontents
\newpage

\section{Introduction}

Pure states: $\mathcal{S} $
Statistical ensemble: $\mathcal{D} $
Mixed state: $\mathcal{M} $
Evolution: $\mathcal{E}$
Fast time scale: $\mathcal{P}$

\section{From projections to unitary evolution}

In this section we will assume as given the state space of non-relativistic quantum mechanics and the rules for projections. We will then recover unitary evolution (i.e. the Schroedinger equation) as the quasi-static deterministic evolution given by continuous projections. To do so, we will introduce a different mathematical notation which will make explicit the difference between pure states, statistical distributions and mixed state, which helps to dispel some confusions. This notation will also prepare the stage for the next section, where we will see that that the same mathematical framework applies to all scientific theories.

We start by having a set of pure states $\mathcal{S}$, which are typically represented in quantum mechanics as directions in a Hilbert space,\footnote{As long as one is able to define projections, other representations can be equivalently used.} so we'll write $\psi \in \mathcal{S}$ meaning $\psi$ is a state for our system. These elements correspond to the possible configurations of a single instance of our system at a given time.

We then define the set of statistical ensembles $\mathcal{D}$, which represent the (classical) statistical distributions of multiple copies of our system independently prepared. Each element of $\mathcal{d} \in \mathcal{D}$ can be represented as a normalized distribution $\mathcal{d}: \mathcal{S} \to \mathbb{R}$ where $\mathcal{d}(\psi)$ represents the probability to find an instance of the distribution in the state $\psi$. For example, if $\mathcal{d}(\psi) = \frac{1}{2}$ then $50\%$ of the times a system prepared according to the statistical ensemble $\mathcal{d}$ will be found in the state $\psi$ . We can also have statistical ensembles prepared in such a way that all instances are prepared in the same pure state. We therefore define $\iota : \mathcal{S} \to \mathcal{D}$ as the statistical ensemble where all instances are prepared in the same way: if $\mathcal{d}=\iota(\psi)$ then $\mathcal{d}(\phi)$ is equal to 1 if $\psi = \phi$ and is equal to 0 otherwise. Mathematically, $\iota$ is an injection of the pure states into the space of statistical distributions. The space $\mathcal{D}$ is a real vector space, as we can combine probability distributions through linear combinations, and the set $\iota(\mathcal{S})$ forms a basis of that space, as any distribution can be constructed from pure states.

Note that $\mathcal{S}$ and $\mathcal{D}$ are two different vector spaces, both mathematically and physically. Suppose we are studying a spin $1/2$ system. Let $z^+, z^- \in \mathcal{S}$ be the two states with spin aligned and anti-aligned with respect to the $z$ direction. We have:
\begin{equation}
\begin{aligned}
\ket{y^+} &= \frac{1}{\sqrt{2}} (\ket{z^+} + i \ket{z^-}) \\
\mathcal{d}&= \frac{1}{2} (\iota(z^+) + \iota(z^-))
\end{aligned}
\end{equation}
The first is a complex linear combination in $\mathcal{S}$ and represents a quantum superposition, which simply returns another pure state. In the example, it returns spin prepared in a different direction. The second is a real linear combination in $\mathcal{D}$ and represents a statistical mixture. In the example, it returns a statistical ensemble that half the time is prepared in one direction and half the time in the opposite direction. The first type of combination is specific to quantum states. The second type of combination applies to all statistical distributions.

In quantum mechanics, not all statistical ensembles are distinguishable from each other. For example, in the spin $1/2$ case, the distribution $\mathcal{d}_1$ defined as $\mathcal{d}_1(z^+)=\mathcal{d}_1(z^-)=\frac{1}{2}$ is physically equivalent to $\mathcal{d}_2$ defined as $\mathcal{d}_2(x^+)=\mathcal{d}_2(x^-)=\frac{1}{2}$. Therefore we define the set of mixed states $\mathcal{M}$ as the set of statistical ensembles that are physically different, which is usually represented by a density matrix. Mathematically, we can also represent physical equivalence between statistical ensemble with the equivalence operator $\mathcal{d}_1 \equiv \mathcal{d}_2$ and the set of mixed states $\mathcal{M} = \mathcal{D}_{/_\equiv}$ will be the quotient space.

We can now introduce the space of projections $\mathcal{P}$ as those processes $\mathcal{p} : \mathcal{D} \to \mathcal{D}$ that take a statistical ensemble and return another statistical ensemble according to the rules of quantum mechanics projections. That is, given a Hermitian operator $O$ over $\mathcal{S}$, we find the set of eigenstates $\{\phi_i\} \subset \mathcal{S}$ for $O$ and we construct the projection $\mathcal{p}_O$ such that:
\begin{equation}\label{transition_probability}
\mathcal{p}_O(\iota(\psi))(\phi) = 
\begin{cases}
\frac{\braket{\psi | \phi} \braket{\phi|\psi}}{\braket{\psi|\psi}} &\quad\phi \in \{\phi_i\}\\
0 &\quad\phi \notin \{\phi_i\}\\
\end{cases}
\end{equation}
If we prepare a statistical ensemble $\iota(\psi)$, where all instances are in state $\psi$, we then apply the process $\mathcal{p}_O$ we find that we cannot be in a state that is not an eigenstate of $O$, and the probability of being in a particular eigenstate is given by the inner product. Since $\iota(\mathcal{S})$ is a basis for $\mathcal{D}$ and the projections are linear, each $\mathcal{p}_O$ is well defined over the whole $\mathcal{D}$.

All we said so far are standard concepts in quantum mechanics re-expressed in a new notation which allows us to clarify an important detail. The processes $\mathcal{P}$ are defined on the statistical ensembles $\mathcal{D}$, not on the pure states $\mathcal{S}$. They are, in general, non-deterministic processes and therefore do not take a pure state to a single well-defined pure state. These processes are projections because they are:
\begin{description}
	\item [linear operations] the output for an ensemble is the same as an ensemble of the outputs for the elements, i.e. $\mathcal{p}(a_1\mathcal{d}_1 + a_2\mathcal{d}_2) = a_1\mathcal{p}(\mathcal{d}_1) + a_2\mathcal{p}(\mathcal{d}_2)$
	\item [idempotent] if applied twice return the same result, i.e. $\mathcal{p}(\mathcal{p}(\mathcal{d})) = \mathcal{p}(\mathcal{d})$
\end{description}
The linearity of these operations, then, is the one defined on statistical ensembles: it is the linearity of classical probability, not of superposition. The fact that these processes are idempotent simply means that, if we run the process again on the output, we get the same output. That is, all the outputs are equilibria. Mathematically, if $\psi$ is an eigenstate of $O$, it is an equilibrium for the process since $\mathcal{p}_O(\iota(\psi))=\iota(\psi)$: every time we start with $\psi$ we always get $\psi$. Therefore, \textbf{we will take the projections to represent non-deterministic processes that always reach equilibria in a very small time $dt$, faster than our description allows}. Similarly to thermodynamics processes, we do not claim a description in terms of our state space $\mathcal{S}$ is valid during that process, only that it is a valid description for the final equilibrium.

Conversely, note that for any pure state $\psi \in \mathcal{S}$ we can find an appropriate Hermitian operator $O$ such that $\psi$ is an eigenstate of $O$ and therefore any pure state is an equilibrium of some process $\mathcal{p} \in \mathcal{P}$. That is, not only the equilibria of every process in $\mathcal{P}$ are elements of $\mathcal{S}$, but all elements of $\mathcal{S}$ are equilibria for at least one process in $\mathcal{P}$: \textbf{all states are equilibrium states of the projections}.

If we accept this premise, then the natural thing to do is to write the conditions for a quasi-static deterministic evolution. That is, we assume that at each moment the system is in equilibrium and we change the state just a little so that each pure state is mapped to a single well-defined pure state. That is, if $\psi_{t}$ and $\psi_{t+dt}$ are two states infinitesimally close in time and $\mathcal{p}_{t+dt}$ is the process that between $[t, t+dt]$ transitions one into the other, then every time $\psi_{t}$ is the input of that process $\psi_{t+dt}$ is its output. Mathematically:
\begin{equation}
\begin{aligned}
\mathcal{p}_{t+dt}(\iota(\psi_{t}))&=\iota(\psi_{t+dt}) \\
\end{aligned}
\end{equation}
Applying the definitions we find:
\begin{equation}
\begin{aligned}
\mathcal{p}_{t+dt}(\iota(\psi_{t}))(\psi_{t+dt}) &=
\frac{\braket{\psi_{t} | \psi_{t+dt}} \braket{\psi_{t+dt}|\psi_{t}}}{\braket{\psi_{t}|\psi_{t}}} = 1 \\
\end{aligned}
\end{equation}
This requires the evolution to be unitary and we can use standard techniques from here, which will lead to the Schroedinger equation. In other words, \textbf{the Schroedinger equation is the quasi-static deterministic evolution of a process of continuous projections}. This approach, therefore, proceeds in the opposite way of others: instead of assuming the Schroedinger equation is always valid and deriving projections as a special case, we assume projections are valid approximations of faster scale processes and the Schrodeinger equation is the special quasi-static deterministic case.

We can briefly see why this approach makes intuitive sense. The closer two states are, the greater the chance that one will transition to the other during a projection. Continuous evolution, then, means small deterministic changes. Unitary evolution, in particular, conserves information entropy, which is also the hallmark of deterministic and reversible evolution. On the other hand, a projection can increase entropy, which is the hallmark of a transition that went out of equilibrium and is not reversible. Since any two unitary vectors can be connected by a unitary operation (i.e. a rotation), it makes sense that all states are equilibria since they can all be connected by a deterministic and reversible process, and it also makes sense that they all carry the same entropy. We also note that in quantum field theory, to calculate cross-sections and particle decays, one puts the initial states at minus infinity and the final states at plus infinity for processes that, in reality, take a tiny fraction of a second. To us, this does sound like a quasi-static approximation. Finally, while the approach presented here does not solve the so called ``measurement problem'', it does however give us an understanding of why we have the problem to begin: our states are really just the equilibrium states of faster scale processes and those cannot, in general, be described by quasi-static ones.

%The connection to measurement is simply that, as we will see later in more detail, such processes are always required if we want to measure a quantity, even in classical mechanics or thermodynamics. For example, if we want to measure the temperature of a gas it needs to be in equilibrium for long enough for us to conduct the measurement. If we are measuring the position of a ball, we need to integrate over a (small) period of time to integrate the giggling created by the random collisions of air particles. That is: if we want to measure the vertical spin component of an electron, something needs to make sure that we actually have and electron and its vertical component is well defined. So we need a process whose outputs are equilibria with those properties.

\section{States and processes}

The next question, then, is whether what we saw in the previous section is an actual model for physical reality or is just a coincidental. We will, once again, proceed in a different way than usual. Instead of justifying the mathematical structure for quantum mechanics in particular, we will show that a very similar mathematical structure exists in all cases. That is, instead of proceeding from what is reasonable in a particular set of circumstances, we proceed from what is necessary in general and find a conceptual and mathematical framework that is similar to the one we already discussed.

In any physical theory, first one needs to define what is the object under study. We typically call that object \emph{system} and \emph{environment} everything else. Depending on the case, our system may be the solar system, a rocket, a ball, a region of the atmosphere, a misture of gas within a can, a handful of molecules or an electron. We call \emph{state} a particular configuration of that system at a particular time and define $\mathcal{S}$ as the set of all possible configurations.

It should be fairly clear that the state space will depend upon the choice of system. What may be less clear is that it also depends on context. For example, if we are studying the motion of a balloon position and momentum may be enough to identify a state. If, instead, we want to study how the balloon expands or contracts we may need temperature and pressure. \textbf{The same system allows different choices for state, which are contingent upon what properties and situations we choose to study.}

Ultimately, what we want is a set of measurable properties $\xi^a : \mathcal{S} \to \mathbb{R}$ that uniquely identify a state and a law of evolution $\mathcal{E} : \mathcal{S} \times \mathbb{R} \to \mathcal{S}$ such that given a state at one time and a time increment we can predict the state at the future time. For example, if we are studying the motion of a ball the effect of the air bumping chaotically on the surface can be disregarded and therefore position and velocity of the center of mass will suffice. If we are studying the motion of a spec of dust, the air collisions cannot be disregarded and therefore we use a statistical ensemble as the state and, given the initial probability distribution, we will predict the final probability distribution.\footnote{In this case, the measurable properties that identify the state become the probabilities associated to each case.} But if we study the spec of dust in a vacuum, then we can go back to just position and momentum. That is, \textbf{states and deterministic laws are defined interdependently}: within a given situation we define states in a particular way because we have a law they obey and we write the law in a particular way because we have states that obey it.

States and deterministic laws, then, are indeed objective concepts but they are not absolute ones. While it is a matter of fact that those particular properties of that particular system under those circumstances are enough to predict their future values, it is up to us to choose the system, the circumstances and the variables that make for a good combination. Therefore, when discussing state definition, we must be cognizant of that choice or we may unknowingly misconstrue the meaning of our theories. We may believe some results to be features of the system studied while in fact they may be direct consequence of us trying to find states we can manipulate and for which we can write deterministic laws. This is akin to a fisherman noting that there is a minimum size to the fish in the sea, not realizing it is precisely the size of the mesh of his net.

With this is mind, let us explore the following question: could we, in principle, extend the state to exhaust the description of the system? As we saw, that would mean finding a set of circumstances in which the evolution of the entirety of the system is fully determined by the system itself. That is, no part of the system is influenced by the environment in any way. Under those conditions, no interactions with the system are possible. Photons cannot scatter off of it and neither gravitational waves. No way to synchronize clocks between the system and the environment, no way to tell where the system is with respect to the environment. To us in the environment, within those circumstances, the system has vanished. How, then, are we supposed to study it? How are we supposed to experimentally define the properties in those conditions? How are we supposed to validate the law of evolution?

This tells us that there is a tension. On one side we can only say that something is a property of the system if its defined independently of the environment. On the other side, we cannot completely turn off system-environment interaction or there would be no way to operationally define the property. This means that we can never have a complete independent description of the system.\footnote{Ultimately, this is why classical mechanics conceptually fails. See XXX for more discussion.} We need to leave out some ``unstated part'' whose future will depend on the environment. In other words, \textbf{the system always has a boundary with the environment and, through it, is in constant interaction}.

This unavoidable interaction with the environment also comes into play in the definition of a system and its state. For example, we can talk about the position of a ball because we sit on the surface of the earth at around 273 K. If we were on the surface of the sun at around 5778 K, after a short time, we would not be able to talk about a ball. A state and its properties only exist because the system is in some kind of equilibrium with the environment. Therefore even if the unstated part is not part of the state, its interaction with the environment has to be ``suitable'' for the state to be defined in the first place. In other words, \textbf{a physical property of a system is well defined and measurable if and only if the constant interaction with the environment keeps it that way.}\footnote{A possible misconception here is thinking of a process that does not change anything as ``doing nothing''. This is similar to thinking of the vacuum as ``nothing''. If one accepts the idea that the vacuum cannot be ``nothing'' then one always has interaction with the vacuum which cannot be ``doing nothing''.}

Let us define $\mathcal{P}$ as the set of all system-environment interactions. While we cannot characterize their effect on the unstated part, we should characterize their relationship to the state space $\mathcal{S}$. These processes may act on the states non-deterministically since they depend in general on the configuration of the environment. So we define $\mathcal{D}$ as the set of all statistical ensembles over $\mathcal{S}$ and each process $\mathcal{p} \in \mathcal{P}$ is a map $\mathcal{p} : \mathcal{D} \to \mathcal{D}$ that takes a statistical ensemble to another statistical ensemble. Given a state $\mathcal{s} \in \mathcal{S}$ we must find at least a set of circumstances in which it is in equilibrium with the environment for a finite amount of time. That is, if we prepare an ensemble $\iota(\mathcal{s})$ that is made of identically prepared copies of $\mathcal{s}$, we have a process $\mathcal{p} \in \mathcal{P}$ such that $\mathcal{p}(\iota(\mathcal{s})) = \iota(\mathcal{s})$.

Every process that has equilibria is characterized by a time scale needed to reach equilibrium. That time will set the maximum time resolution we can use with that choice of states and processes. To go to faster time scales we inevitably need to change both states and processes. For example, as we change the volume of a gas, the region against the moving wall will be compressed sooner than the opposite region, but if we wait a short while we will again obtain a uniform pressure. If we want to study that transition, we may conceptually divide the system into small volumes and the pressure is replaced by a pressure field. Still, we need to assume each small volume to have enough molecules to be in equilibrium. We can go into more detail, and describe the distribution of position and velocities of the molecules. Yet, we need to assume that the molecules themselves are not undergoing chemical changes. We could study the atoms of the molecules, yet we would have to assume that the atoms are not undergoing nuclear changes. We could study the electrons and nucleons of the atoms, yet we would have to assume they are stable under weak nuclear forces. As we move to a finer and finer description, we have faster and faster scale processes that we assume do not change the basic constituents. At each level of the description what is assumed to be in equilibrium changes and what the laws describe change as well. Yet, the overall structure remains the same: we have a timescale $dt$ such that within each interval $[t-dt, t]$ there is a process $\mathcal{p}_{t-dt}$ such that the state $\mathcal{s}_t$ is an equilibrium of $\mathcal{p}_{t-dt}$. The evolution $\mathcal{E}$, then, is quasi-static in terms of the processes in $\mathcal{P}$. That is, \textbf{all deterministic laws are quasi-static evolutions in terms of system-environment interaction processes}.

We have reached, then, the second major point of this article: \textbf{the mathematical structure we used to rearrange quantum mechanics is a general structure that applies to all systems} and as such we call it the \emph{fundamental model of physics}. It is fundamental because it always applies and what changes case by case are the specifics. If we study a ball using classical particle mechanics, for example, the position and momentum of the center of mass will be defined in a short time scale integrating the giggling created by the random collisions of air particles. That averaging process will be the same regardless of the values of position and momentum, all states are equilibria of the same process and therefore $\mathcal{P}$ contains only one element: the identity map. Therefore, while the overall structure still exists, it is mathematically trivial and can be effectively disregarded as we concentrate on the law of evolution exclusively.

In thermodynamics, instead, processes and equilibria play a critical role. The study of the boundary between system and environment becomes the main focus. In this framework, choosing an isolated fixed wall means we picked the process $\mathcal{p}_{U,V} \in \mathcal{P}$ that guarantees the equilibrium to have a fixed energy and volume. The equilibrium state $\mathcal{s} \in \mathcal{S}$ will be associated to other quantities, like pressure, temperature and entropy, that are only defined on the equilibrium state itself. Work and heat are associated with the process, not the state, and that is why, in general, they are not calculable. State evolution occurs because we change the constraints or work parameters, for example we change volume at equal pressure. That is, at each time step we slowly change either the state or the process and let the system reach a new equilibrium.

In quantum mechanics we have elements of both. Like in classical mechanics we have a state space and a law of evolution, given by the choice of an Hamiltonian, that is deterministic and reversible. Like in thermodynamics, we have states that are equilibria of non-deterministic processes, of which we are not able to give a more fundamental description. Quantities are only well defined at equilibrium and the law of evolution is the quasi-static evolution where we assume the system changes slowly and we are always at equilibrium. What is interesting is that \textbf{in quantum mechanics what drives the deterministic evolution are the same processes that drive the equilibrium}.

In this framework all processes in $\mathcal{P}$ always represent abrupt transitions to a new equilibrium. In thermodynamics they could represent the crystallization of a supersaturated solution or the ignition of a mixture of oxygen and hydrogen to form water. In quantum mechanics they represent the absorption of an incoming photon by an hydrogen atom or the decay of a muon into an electron and two neutrinos. In all these cases we start from a state of (possibly metastable) equilibrium and we end with a new equilibrium which is not, in general, described in terms of the initial state space. For example, water vapor is not a mixture of hydrogen and oxygen gases and a muon is not a collection of three particles. That is, in general the system passes through configurations that are not suitably described by the initial or final state space. We do not believe the similar conceptual structure to be simply a formal analogy but to represent analogue physical mechanisms. Quantum projections happen for the same reasons supersaturated solution crystallize: an abrupt change in external or internal physical conditions.

Why, then, are measurements in quantum mechanics typically associated with projections? Let us think of what a particle detector has to do: it has to take the very small signal of a particle coming through and amplify it to a macroscopic level. In the case the particle itself is absorbed, or destroyed in any way, then are already in the realm of fast processes. In the other cases, the particle has to initiate a chain reaction that has to transfer a lot more energy than the particle itself has. The detector, then, has to be prepared in a state sufficiently stable to not change if no particle interacts with it but unstable enough that a small perturbation makes it transition to a lower energy. That is, the detector has to be prepared in a metastable state and the particle interaction has to make it transition to a more stable state. But a transition from a metastable state to a more stable state is exactly what our projections are supposed to represent. %https://www.researchgate.net/publication/280330121_A_Heuristic_Approach_to_the_Quantum_Measurement_Problem_How_to_Distinguish_Particle_Detectors_from_Ordinary_Objects

We can sum up this section in the following points:
\begin{itemize}
	\item the mathematical structure of quantum mechanics (i.e. projections and law of evolution) is the natural structure that all physical theories have
	\item projections represent processes with equilibria that happen at a faster scale and in general not describable in terms of said description
	\item measurements are simply physical processes
\end{itemize}
The advantage of this perspective is that no ad-hoc concepts are needed. In future works we will further characterize this model providing even more constraints to physical theories, which may clarify what other elements of processes and laws are basic requirements and which are particular to classical or quantum mechanics.

\section{???}

Now that we have presented the fundamental model and saw how quantum mechanics is simply a specific instance, we have a broader framework to gain insights.

New physics should be found for processes that cannot be described by either unitary evolution or projection. Out of equilibrium. But the issue is that:

Quantum states may not provide the ultimate description for states. We will need states for the out of equilibrium. But the issues is that:

At some point, time cannot be continuous. The idea is that the scale of the process is no longer faster than that of the evolution. We will need new mathematical tools.

\bibliography{bibliography}

\end{document}

### 2019-03-16 Articles/2019-ForQuantumOptics/article.tex
Review with Christine and first pass at last section

@@
 the Schroedinger equation) as the quasi-static deterministic evolution given by continuous projections.
-To do so, we will introduce a different mathematical notation which will make explicit the difference between pure states, statistical distributions and mixed state, which helps to dispel some confusions.
-This notation will also prepare the stage for the next section, where we will see that that the same mathematical framework applies to all scientific theories.
+To do so, we will introduce a different mathematical notation which will make explicit the difference between pure states, statistical distributions and mixed states, to help dispel potential confusion.
+This notation will also prepare the stage for the next section, where we will see that the same mathematical framework applies to all scientific theories.
 
-We start by having a set of pure states $\mathcal{S}$, which are typically represented in quantum mechanics as directions in a Hilbert space,\footnote{As long as one is able to define projections, other representations can be equivalently used.} so we'll write $\psi \in \mathcal{S}$ meaning $\psi$ is a state for our system.
-These elements correspond to the possible configurations of a single instance of our system at a given time.
+We start by having a set of pure states $\mathcal{S}$, which are typically represented in quantum mechanics as directions in a Hilbert space,\footnote{As long as one is able to define projections, other representations can be equivalently used.} so we will write $\psi \in \mathcal{S}$ meaning $\psi$ is a state for our system.
+The elements in $\mathcal{S}$ correspond to the possible configurations of a single instance of our system at a given time.
 
 We then define the set of statistical ensembles $\mathcal{D}$, which represent the (classical) statistical distributions of multiple copies of our system independently prepared.
-Each element of $\mathcal{d} \in \mathcal{D}$ can be represented as a normalized distribution $\mathcal{d}: \mathcal{S} \to \mathbb{R}$ where $\mathcal{d}(\psi)$ represents the probability to find an instance of the distribution in the state $\psi$.
-For example, if $\mathcal{d}(\psi) = \frac{1}{2}$ then $50\%$ of the times a system prepared according to the statistical ensemble $\mathcal{d}$ will be found in the state $\psi$ .
+Each element $\mathcal{d} \in \mathcal{D}$ can be represented as a normalized distribution $\mathcal{d}: \mathcal{S} \to \mathbb{R}$ where $\mathcal{d}(\psi)$ represents the probability to find an instance of the distribution in the state $\psi$.
+For example, if $\mathcal{d}(\psi) = \frac{1}{2}$ then $50\%$ of the times a system prepared according to the statistical ensemble $\mathcal{d}$ will be found in the state $\psi$.
 We can also have statistical ensembles prepared in such a way that all instances are prepared in the same pure state.
-We therefore define $\iota : \mathcal{S} \to \mathcal{D}$ as the statistical ensemble where all instances are prepared in the same way: if $\mathcal{d}=\iota(\psi)$ then $\mathcal{d}(\phi)$ is equal to 1 if $\psi = \phi$ and is equal to 0 otherwise.
+We therefore define $\iota : \mathcal{S} \to \mathcal{D}$ as the function that returns a statistical ensemble where all instances are prepared in the same way: if $\mathcal{d}=\iota(\psi)$ then $\mathcal{d}(\phi)$ is equal to 1 if $\phi = \psi$ and is equal to 0 otherwise.
 Mathematically, $\iota$ is an injection of the pure states into the space of statistical distributions.
@@
 The second is a real linear combination in $\mathcal{D}$ and represents a statistical mixture.
-In the example, it returns a statistical ensemble that half the time is prepared in one direction and half the time in the opposite direction.
+In the example, it returns a statistical ensemble that is prepared half the time in one direction and half the time in the opposite direction.
 The first type of combination is specific to quantum states.
@@
 Therefore we define the set of mixed states $\mathcal{M}$ as the set of statistical ensembles that are physically different, which is usually represented by a density matrix.
-Mathematically, we can also represent physical equivalence between statistical ensemble with the equivalence operator $\mathcal{d}_1 \equiv \mathcal{d}_2$ and the set of mixed states $\mathcal{M} = \mathcal{D}_{/_\equiv}$ will be the quotient space.
+Mathematically, we can also represent physical equivalence between statistical ensembles with the equivalence operator $\mathcal{d}_1 \equiv \mathcal{d}_2$ and the set of mixed states $\mathcal{M} = \mathcal{D}_{/_\equiv}$ will be the quotient space.
 
-We can now introduce the space of projections $\mathcal{P}$ as those processes $\mathcal{p} : \mathcal{D} \to \mathcal{D}$ that take a statistical ensemble and return another statistical ensemble according to the rules of quantum mechanics projections.
+We can now introduce the space of projections $\mathcal{P}$ as those processes $\mathcal{p} : \mathcal{D} \to \mathcal{D}$ that take a statistical ensemble and return another statistical ensemble according to the rules of projections in quantum mechanics.
 That is, given a Hermitian operator $O$ over $\mathcal{S}$, we find the set of eigenstates $\{\phi_i\} \subset \mathcal{S}$ for $O$ and we construct the projection $\mathcal{p}_O$ such that:
@@
 \end{equation}
-If we prepare a statistical ensemble $\iota(\psi)$, where all instances are in state $\psi$, we then apply the process $\mathcal{p}_O$ we find that we cannot be in a state that is not an eigenstate of $O$, and the probability of being in a particular eigenstate is given by the inner product.
+If we prepare a statistical ensemble $\iota(\psi)$, where all instances are in state $\psi$, and we subsequently apply the process $\mathcal{p}_O$, then we will find that we cannot be in a state that is not an eigenstate of $O$, and the probability of being in a particular eigenstate is given by the inner product.
 Since $\iota(\mathcal{S})$ is a basis for $\mathcal{D}$ and the projections are linear, each $\mathcal{p}_O$ is well defined over the whole $\mathcal{D}$.
 
-All we said so far are standard concepts in quantum mechanics re-expressed in a new notation which allows us to clarify an important detail.
-The processes $\mathcal{P}$ are defined on the statistical ensembles $\mathcal{D}$, not on the pure states $\mathcal{S}$.
+All we presented so far are standard concepts in quantum mechanics re-expressed in a new notation which allows us to clarify an important detail: the processes $\mathcal{P}$ are defined on the statistical ensembles $\mathcal{D}$, not on the pure states $\mathcal{S}$.
 They are, in general, non-deterministic processes and therefore do not take a pure state to a single well-defined pure state.
@@
 That is, all the outputs are equilibria.
-Mathematically, if $\psi$ is an eigenstate of $O$, it is an equilibrium for the process since $\mathcal{p}_O(\iota(\psi))=\iota(\psi)$: every time we start with $\psi$ we always get $\psi$.
+Mathematically, if $\psi$ is an eigenstate of $O$, it is an equilibrium for the process since $\mathcal{p}_O(\iota(\psi))=\iota(\psi)$.
+Every time we start with $\psi$ we always get $\psi$.
 Therefore, \textbf{we will take the projections to represent non-deterministic processes that always reach equilibria in a very small time $dt$, faster than our description allows}.
-Similarly to thermodynamics processes, we do not claim a description in terms of our state space $\mathcal{S}$ is valid during that process, only that it is a valid description for the final equilibrium.
+Similarly to thermodynamics processes, we do not claim that a description in terms of our state space $\mathcal{S}$ is valid during that process, only that it is a valid description for the final equilibrium.
 
 Conversely, note that for any pure state $\psi \in \mathcal{S}$ we can find an appropriate Hermitian operator $O$ such that $\psi$ is an eigenstate of $O$ and therefore any pure state is an equilibrium of some process $\mathcal{p} \in \mathcal{P}$.
-That is, not only the equilibria of every process in $\mathcal{P}$ are elements of $\mathcal{S}$, but all elements of $\mathcal{S}$ are equilibria for at least one process in $\mathcal{P}$: \textbf{all states are equilibrium states of the projections}.
+That is, not only are the equilibria of every process in $\mathcal{P}$ elements of $\mathcal{S}$, but all elements of $\mathcal{S}$ are equilibria for at least one process in $\mathcal{P}$.
+\textbf{All states are equilibrium states of the projections}.
 
@@
 In other words, \textbf{the Schroedinger equation is the quasi-static deterministic evolution of a process of continuous projections}.
-This approach, therefore, proceeds in the opposite way of others: instead of assuming the Schroedinger equation is always valid and deriving projections as a special case, we assume projections are valid approximations of faster scale processes and the Schrodeinger equation is the special quasi-static deterministic case.
+This approach, therefore, proceeds in the opposite way of others: instead of assuming the Schroedinger equation is always valid and deriving projections as a special case, we assume projections are valid approximations of faster scale processes and the Schroedinger equation is the special quasi-static deterministic case.
 
 We can briefly see why this approach makes intuitive sense.
-The closer two states are, the greater the chance that one will transition to the other during a projection.
+The closer two states are in the Hilbert space, the greater the chance that one will transition to the other during a projection.
 Continuous evolution, then, means small deterministic changes.
-Unitary evolution, in particular, conserves information entropy, which is also the hallmark of deterministic and reversible evolution.
+Unitary evolution, in particular, conserves entropy, which is also the hallmark of deterministic and reversible evolution.
 On the other hand, a projection can increase entropy, which is the hallmark of a transition that went out of equilibrium and is not reversible.
-Since any two unitary vectors can be connected by a unitary operation (i.e.
-a rotation), it makes sense that all states are equilibria since they can all be connected by a deterministic and reversible process, and it also makes sense that they all carry the same entropy.
+Since any two unitary vectors can be connected by a unitary operation (i.e.~a rotation), it makes sense that all states are equilibria since they can all be connected by a deterministic and reversible process and that they all carry the same entropy.
 We also note that in quantum field theory, to calculate cross-sections and particle decays, one puts the initial states at minus infinity and the final states at plus infinity for processes that, in reality, take a tiny fraction of a second.
-To us, this does sound like a quasi-static approximation.
-Finally, while the approach presented here does not solve the so called ``measurement problem'', it does however give us an understanding of why we have the problem to begin: our states are really just the equilibrium states of faster scale processes and those cannot, in general, be described by quasi-static ones.
+This is reminiscent of a quasi-static approximation.
+Finally, while the approach presented here does not solve the so called ``measurement problem'', it does however give us an understanding of why we have the problem to begin with: our states are really just the equilibrium states of faster scale processes and those cannot, in general, be described by quasi-static ones.
 
@@
 
-The next question, then, is whether what we saw in the previous section is an actual model for physical reality or is just a coincidental.
+The next question, then, is whether what we saw in the previous section is an actual model for physical reality.
 We will, once again, proceed in a different way than usual.
-Instead of justifying the mathematical structure for quantum mechanics in particular, we will show that a very similar mathematical structure exists in all cases.
-That is, instead of proceeding from what is reasonable in a particular set of circumstances, we proceed from what is necessary in general and find a conceptual and mathematical framework that is similar to the one we already discussed.
+Instead of justifying the mathematical structure for quantum mechanics in particular, we will show that a very similar mathematical structure exists when studying other systems.
+That is, instead of proceeding from what is reasonable in a particular set of circumstances, we proceed from what is necessary in general and find a conceptual and mathematical framework that is similar to the one already discussed.
 
@@
 What may be less clear is that it also depends on context.
-For example, if we are studying the motion of a balloon position and momentum may be enough to identify a state.
+For example, if we are studying the motion of a balloon, position and momentum may be enough to identify a state.
 If, instead, we want to study how the balloon expands or contracts we may need temperature and pressure.
@@
 
-Ultimately, what we want is a set of measurable properties $\xi^a : \mathcal{S} \to \mathbb{R}$ that uniquely identify a state and a law of evolution $\mathcal{E} : \mathcal{S} \times \mathbb{R} \to \mathcal{S}$ such that given a state at one time and a time increment we can predict the state at the future time.
+Ultimately, what we want is a set of measurable properties $\xi^a : \mathcal{S} \to \mathbb{R}$ that uniquely identify a state, and a law of evolution $\mathcal{E} : \mathcal{S} \times \mathbb{R} \to \mathcal{S}$ such that given a state at one time and a time increment we can predict the state at the future time.
 For example, if we are studying the motion of a ball the effect of the air bumping chaotically on the surface can be disregarded and therefore position and velocity of the center of mass will suffice.
-If we are studying the motion of a spec of dust, the air collisions cannot be disregarded and therefore we use a statistical ensemble as the state and, given the initial probability distribution, we will predict the final probability distribution.\footnote{In this case, the measurable properties that identify the state become the probabilities associated to each case.} But if we study the spec of dust in a vacuum, then we can go back to just position and momentum.
-That is, \textbf{states and deterministic laws are defined interdependently}: within a given situation we define states in a particular way because we have a law they obey and we write the law in a particular way because we have states that obey it.
+If we are studying the motion of a speck of dust, the air collisions cannot be disregarded and therefore we use a statistical ensemble as the state and, given the initial probability distribution, we predict the final probability distribution.\footnote{In this case, the measurable properties that identify the state become the probabilities associated to each case.} But if we study the speck of dust in a vacuum, then we can go back to just position and momentum.
+That is, \textbf{states and deterministic laws are defined interdependently}.
+Within a given situation we define states in a particular way because we have a law they obey, and we write the law in a particular way because we have states that obey it.
 
 States and deterministic laws, then, are indeed objective concepts but they are not absolute ones.
-While it is a matter of fact that those particular properties of that particular system under those circumstances are enough to predict their future values, it is up to us to choose the system, the circumstances and the variables that make for a good combination.
+While it is a matter of fact that those particular properties of that particular system under those circumstances are enough to predict their future values, it is up to us to choose the system, the circumstances and the variables that constitute an appropriate combination.
 Therefore, when discussing state definition, we must be cognizant of that choice or we may unknowingly misconstrue the meaning of our theories.
-We may believe some results to be features of the system studied while in fact they may be direct consequence of us trying to find states we can manipulate and for which we can write deterministic laws.
+We may believe some results to be features of the system studied while in fact they may be direct consequences of our trying to find states we can manipulate and for which we can write deterministic laws.
 This is akin to a fisherman noting that there is a minimum size to the fish in the sea, not realizing it is precisely the size of the mesh of his net.
@@
 Under those conditions, no interactions with the system are possible.
-Photons cannot scatter off of it and neither gravitational waves.
-No way to synchronize clocks between the system and the environment, no way to tell where the system is with respect to the environment.
+Photons cannot scatter off of it, nor gravitational waves; no way to synchronize clocks between the system and the environment, no way to tell where the system is with respect to the environment.
 To us in the environment, within those circumstances, the system has vanished.
@@
 This tells us that there is a tension.
-On one side we can only say that something is a property of the system if its defined independently of the environment.
+On one side we can only say that something is a property of the system if it is defined independently of the environment.
 On the other side, we cannot completely turn off system-environment interaction or there would be no way to operationally define the property.
@@
 Therefore even if the unstated part is not part of the state, its interaction with the environment has to be ``suitable'' for the state to be defined in the first place.
-In other words, \textbf{a physical property of a system is well defined and measurable if and only if the constant interaction with the environment keeps it that way.}\footnote{A possible misconception here is thinking of a process that does not change anything as ``doing nothing''.
-This is similar to thinking of the vacuum as ``nothing''.
-If one accepts the idea that the vacuum cannot be ``nothing'' then one always has interaction with the vacuum which cannot be ``doing nothing''.}
+In other words, \textbf{a physical property of a system is well defined and measurable if and only if the constant interaction with the environment keeps it that way.}\footnote{A possible misconception here is thinking of a process that does not change anything as ``doing nothing''.}
 
@@
 So we define $\mathcal{D}$ as the set of all statistical ensembles over $\mathcal{S}$ and each process $\mathcal{p} \in \mathcal{P}$ is a map $\mathcal{p} : \mathcal{D} \to \mathcal{D}$ that takes a statistical ensemble to another statistical ensemble.
-Given a state $\mathcal{s} \in \mathcal{S}$ we must find at least a set of circumstances in which it is in equilibrium with the environment for a finite amount of time.
+Given a state $\mathcal{s} \in \mathcal{S}$ we must find at least one set of circumstances in which it is in equilibrium with the environment for a finite amount of time.
 That is, if we prepare an ensemble $\iota(\mathcal{s})$ that is made of identically prepared copies of $\mathcal{s}$, we have a process $\mathcal{p} \in \mathcal{P}$ such that $\mathcal{p}(\iota(\mathcal{s})) = \iota(\mathcal{s})$.
@@
 Yet, we need to assume that the molecules themselves are not undergoing chemical changes.
-We could study the atoms of the molecules, yet we would have to assume that the atoms are not undergoing nuclear changes.
+We could study the atoms of the molecules, yet we would have to assume that the atoms are stable under nuclear fission.
 We could study the electrons and nucleons of the atoms, yet we would have to assume they are stable under weak nuclear forces.
 As we move to a finer and finer description, we have faster and faster scale processes that we assume do not change the basic constituents.
-At each level of the description what is assumed to be in equilibrium changes and what the laws describe change as well.
-Yet, the overall structure remains the same: we have a timescale $dt$ such that within each interval $[t-dt, t]$ there is a process $\mathcal{p}_{t-dt}$ such that the state $\mathcal{s}_t$ is an equilibrium of $\mathcal{p}_{t-dt}$.
+At each level of the description what is assumed to be in equilibrium changes and what the laws describe changes as well.
+Yet, the overall structure remains the same: we have a timescale $dt$ such that within each interval $[t-dt, t]$ there is a process $\mathcal{p}_t$ such that the state $\mathcal{s}_t$ is an equilibrium of $\mathcal{p}_t$.
 The evolution $\mathcal{E}$, then, is quasi-static in terms of the processes in $\mathcal{P}$.
@@
 It is fundamental because it always applies and what changes case by case are the specifics.
-If we study a ball using classical particle mechanics, for example, the position and momentum of the center of mass will be defined in a short time scale integrating the giggling created by the random collisions of air particles.
-That averaging process will be the same regardless of the values of position and momentum, all states are equilibria of the same process and therefore $\mathcal{P}$ contains only one element: the identity map.
+If we study a ball using classical particle mechanics, for example, the position and momentum of the center of mass will be defined in a short time scale integrating the jiggling created by the random collisions of air particles.
+That averaging process will be the same regardless of the values of position and momentum; all states are equilibria of the same process and therefore $\mathcal{P}$ contains only one element: the identity map.
 Therefore, while the overall structure still exists, it is mathematically trivial and can be effectively disregarded as we concentrate on the law of evolution exclusively.
@@
 In quantum mechanics we have elements of both.
-Like in classical mechanics we have a state space and a law of evolution, given by the choice of an Hamiltonian, that is deterministic and reversible.
+Like in classical mechanics we have a state space and a law of evolution, given by the choice of a Hamiltonian, that is deterministic and reversible.
 Like in thermodynamics, we have states that are equilibria of non-deterministic processes, of which we are not able to give a more fundamental description.
@@
 In thermodynamics they could represent the crystallization of a supersaturated solution or the ignition of a mixture of oxygen and hydrogen to form water.
-In quantum mechanics they represent the absorption of an incoming photon by an hydrogen atom or the decay of a muon into an electron and two neutrinos.
+In quantum mechanics they represent the absorption of an incoming photon by a hydrogen atom or the decay of a muon into an electron and two neutrinos.
 In all these cases we start from a state of (possibly metastable) equilibrium and we end with a new equilibrium which is not, in general, described in terms of the initial state space.
@@
 We do not believe the similar conceptual structure to be simply a formal analogy but to represent analogue physical mechanisms.
-Quantum projections happen for the same reasons supersaturated solution crystallize: an abrupt change in external or internal physical conditions.
+Quantum projections happen for the same reasons supersaturated solutions crystallize: an abrupt change in external or internal physical conditions.
 
@@
 Let us think of what a particle detector has to do: it has to take the very small signal of a particle coming through and amplify it to a macroscopic level.
-In the case the particle itself is absorbed, or destroyed in any way, then are already in the realm of fast processes.
-In the other cases, the particle has to initiate a chain reaction that has to transfer a lot more energy than the particle itself has.
+In the case the particle itself is absorbed, or destroyed in any way, then we are already in the realm of fast processes.
+In the other cases, the particle has to initiate a chain reaction that has to transfers energy stored in the detector system to the readout.
 The detector, then, has to be prepared in a state sufficiently stable to not change if no particle interacts with it but unstable enough that a small perturbation makes it transition to a lower energy.
@@
 	\item the mathematical structure of quantum mechanics (i.e. projections and law of evolution) is the natural structure that all physical theories have
-	\item projections represent processes with equilibria that happen at a faster scale and in general not describable in terms of said description
+	\item projections represent processes with equilibria that happen at a faster scale 
 	\item measurements are simply physical processes
@@
 The advantage of this perspective is that no ad-hoc concepts are needed.
-In future works we will further characterize this model providing even more constraints to physical theories, which may clarify what other elements of processes and laws are basic requirements and which are particular to classical or quantum mechanics.
+In future works we will further characterize this model providing even more constraints to physical theories, which may clarify which other elements of processes and laws are basic requirements and which are particular to classical or quantum mechanics.
 
@@
 
-Now that we have presented the fundamental model and saw how quantum mechanics is simply a specific instance, we have a broader framework to gain insights.
+If we accept the line of thinking present in the previous two sections, we can try and draw some conclusions.
+These will be necessary qualitative as much further work would be needed to make any quantitative prediction.
 
-New physics should be found for processes that cannot be described by either unitary evolution or projection. Out of equilibrium. But the issue is that:
+The first main consequence is that we should be able to find processes that are neither described by unitary evolution nor by the projection.
+These would correspond to processes for which the out-of-equilibrium dynamics becomes relevant so the quasi-static evolution is not a good approximation and neither is the instantaneous transition.
+Even if we do not have a theory for those regimes, there aren't so many ways to achieve it so we can, at least conceptually, create a few categories.
 
-Quantum states may not provide the ultimate description for states. We will need states for the out of equilibrium. But the issues is that:
+For example, slow changing conditions are unlikely to reveal new dynamics.
+As long as we change things adiabatically, the quasi-static approximation will hold.
+If we were to have a process that end-to-end can be modeled by a projection, but was implemented as a slow transition, the quasi-static approximation holds and no new phenomena would be osserved.
 
-At some point, time cannot be continuous. The idea is that the scale of the process is no longer faster than that of the evolution. We will need new mathematical tools.
+The main issue is that we have to catch the system out of equilibrium.
+In principle, we can imagine preparing a system, starting a process $\mathcal{p}_1$ and, while in transition, switch to process $\mathcal{p}_2$.
+At that point, the final state may have characteristics that are different from having run the two process sequentially and that would be evidence of new physics.
+This may correspond to having an atom absorb a photon while it was already in the process of emitting one.
+Or it may correspond to measure the horizontal spin of a particle while we where in the process of measuring the vertical one.
+Whether anything along these lines is actually practicable is unclear, but it's one category.
+
+Another venue would be to find a process that under some conditions is well described by unitary evolution while in others by a projection.
+We may find that at the boundary of the transition the process will be described by neither.
+This may correspond, for example, to those cases where an energy threshold has to be passed to see particle productions or particle absorption.
+It is not clear what kind of precision would be required to elicit a different response, but it's another category.
+
+The second main consequence is that to get a more fundamental theory we have to change the state space.
+As quantum states only characterize the equilibria for the projections, we need a different state space for the out-of-equilibrium states.
+This suggests that developing the right physical model for what happens during a projection is not a simple matter of finding a suitable Hamiltonian or a non-linear law of evolution.
+We may have to substitute complex Hilbert spaces with something else.
+Given that we have no consensus on why those spaces are needed in the first place, and on whether all the elements are actually physical, it is unclear what we would substitute them with.
+
+But it is actually worse than that.
+If we want a theory that works at the smallest time resolution, it is obvious that our $\mathcal{P}$ would correspond to those processes.
+Those will correspond to the basic building blocks with no further refinement possible.
+But in those case the temporal scale of the time evolution is the same scale of the processes: we cannot see time evolution $\mathcal{E}$ as a continuous sequence of processes $\mathcal{p}_t$.
+In fact, there is nothing that would prevent those processes to even be sequential.
+
+In other words, it is possible that we may be very very far from having the correct mathematical tools to be able to characterize those regimes. All
 

### 2019-03-21 Articles/2019-ForQuantumOptics/article.tex
Added introduction and conclusion

@@
 
-\title{TODO}
+\title{Unitary evolution as a quasistatic process of infinitesimal collapses}
 
@@
 \begin{abstract}
-	TODO
+Any state, to be properly defined, has to be an equilibrium of all processes happening at a faster timescale.
+As processes with equilibria are idempotent, they can be expressed as projections in a probability space.
+Assuming a quasistatic evolution, then, means the equilibria of the faster scale processes will evolve slowly, giving rise to continuous and unitary evolution.
+These ideas are enough to get the general form of the projection postulate and the time evolution of quantum mechanics.
+Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
 \end{abstract}
 
-\tableofcontents
-\newpage
+%\tableofcontents
+%\newpage
 
@@
 
-Pure states: $\mathcal{S} $
-Statistical ensemble: $\mathcal{D} $
-Mixed state: $\mathcal{M} $
-Evolution: $\mathcal{E}$
-Fast time scale: $\mathcal{P}$
+In an effort to better understand why the frameworks of classical and quantum mechanics are what they are, we have worked on identifying a minimal set of physical assumptions from which both can be derived.
+This gave us a new perspective that is somewhat unusual among people who fork on fundamental issues in physics.
+The typical view is that the theories are so different that they require a radical different set of starting point, and that classical mechanics in particular should only be viewed as an approximation of the other.
+In this view, the similarities between the theories are due to one being an approximation of the other.
+
+We believe a better view is to think of both classical and quantum mechanics as a specialization of a simpler more fundamental theory.
+This would deal with states and laws of evolution as abstract objects, formalizing what are the \emph{necessary} conditions for them to be operationally defined such that they can be experimentally studied.
+The other physical theories (classical and quantum mechanics, thermodynamics and so on) would specialize those concepts to better define what states we are studying and under what conditions.
+In this view, the similarities between the theories are due to the general concepts of the more fundamental theory.
+This different approach can help us pinpoint what elements are specific to what theory, clarify their realm of applicability and, hopefully, give us ideas where new physics may lie.
+
+In this paper we start clarifying what are the fundamental relationships and, in particular, how they work in the context of quantum mechanics.
+The general idea is that for a state to be experimentally prepared, manipulated and measured, it needs to be ``stable'' for a maybe very short but certainly finite amount of time.
+That is, it must be a, possibly metastable, equilibrium of faster scale processes.
+But a process for which all outputs are equilibria is idempotent (i.e.
+gives the same output no matter how many is applied) and is therefore mathematically modeled by a projection operation on the space of statistical distributions.
+We identify quantum projections as these processes.
+If we want a deterministic law, then, it has to be a quasi-static evolution in terms of those fast scale processes.
+As the change has to be small, this leads to continuous evolution over the state space and unitary evolution in the space of statistical distributions.
+We identify unitary evolution as these processes.
+
+The necessity of states to be stable for a certain minimum timescale applies to all physical theories, and a deterministic law, then, can only be valid \emph{above} that timescale within each theory.
+In other words, fast scale processes define the state space as their equilibria and the minimum scale $dt$ for the law of evolution.
+It will be instructive, then, to see how these concepts apply to classical mechanics and thermodynamics as well.
+This will foster the idea that looking at the foundations of all physical theories, and not on quantum mechanics in isolation, fosters a better general understanding of fundamental concepts.
+
+In section 2 we will first take the basic framework of quantum mechanics and dress it in a slightly different language.
+The purpose is to make the interplay between pure states, statistical distributions of pure states, projections and unitary evolution more clear.
+In that context it will be easier to see how unitary evolution can be seen as a quasi-static process of infinitesimally close projections.
+In section 3 we will see how the same mathematical structure can be justified from more general physical consideration and how it necessarily applies to all physical theories, including classical mechanics in the thermodynamics.
 
@@
 
-\section{???}
+\section{Conclusion}
 
-If we accept the line of thinking present in the previous two sections, we can try and draw some conclusions.
+We have seen that quantum mechanics fits in a more general framework that applies to all other physical theories.
+States are always equilibria of some faster scale process, which in quantum mechanics are modeled with the projection.
+States are chosen to satisfy a law of evolution, which is necessarily a quasi-static process in terms of the faster scale processes, which in quantum mechanics is modeled by the Schroedinger equation.
+This general framework only uses standard physics concepts and puts no special role on the observer except the choice of what processes the system will be subjected to.
+
+Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasi-static.
+If we were to implement a measurement process, then, in a way that the transfer of information between the system and the probe happens in a slow and controlled fashion, this would will not provide a violation to the Schroedinger equation, nor it would tell us that all projections are quasi-static processes.
+A better description of the faster scale processes given by the projections, then, will inevitable need a new state space as the standard quantum states only describes equilibria, and we should expect the faster scale processes to go out of equilibria.
+
+To understand what is the right mathematical framework for those regimes, we are proceeding by first identifying the physical requirements for all mathematical structures already used in physics, starting from the most basic ones like topological spaces (upon which is based all of differential geometry) and $\sigma$-algebras (upon which is based all of measure theory and probability theory).
+This long term effort is centered mainly around the development of an open access book in which these issues can be explored in full detail and with a rigor more typical of mathematical disciplines.
+This rigor is what forces us to spell out the often hidden assumptions we have while constructing physical theories.
+By making explicit under which physically motivated circumstances time can be modeled by real numbers, states can be modeled by classical phase space or by quantum Hilbert spaces, we believe we will be in a better position to understand what happens when we can't.
+
+%\section{???}
+
+%If we accept the line of thinking present in the previous two sections, we can try and draw some conclusions.
 These will be necessary qualitative as much further work would be needed to make any quantitative prediction.
 
-The first main consequence is that we should be able to find processes that are neither described by unitary evolution nor by the projection.
+%The first main conclusion is that we should be able to find processes that are neither described by unitary evolution nor by the projection.
 These would correspond to processes for which the out-of-equilibrium dynamics becomes relevant so the quasi-static evolution is not a good approximation and neither is the instantaneous transition.
-Even if we do not have a theory for those regimes, there aren't so many ways to achieve it so we can, at least conceptually, create a few categories.
+Even if we do not have a theory for those regimes, we can, at least conceptually, discuss a couple of general strategies.
 
-For example, slow changing conditions are unlikely to reveal new dynamics.
-As long as we change things adiabatically, the quasi-static approximation will hold.
-If we were to have a process that end-to-end can be modeled by a projection, but was implemented as a slow transition, the quasi-static approximation holds and no new phenomena would be osserved.
+%For example, we can note that slow changing conditions are unlikely to reveal new dynamics.
+As long as the system evolves continuously, the quasi-static approximation will hold.
+It is also conceivable that, in some conditions, a process that may look like a projection end-to-end is actually well described by a slow transition.
+In those cases the quasi-static approximation evidently holds and no new phenomena would be observed.
+What is the right time scale, though, is not clear to us since ``slow'' in this setting is still ``very fast''.
 
-The main issue is that we have to catch the system out of equilibrium.
-In principle, we can imagine preparing a system, starting a process $\mathcal{p}_1$ and, while in transition, switch to process $\mathcal{p}_2$.
+%The main issue is that we have to catch the system out of equilibrium.
+In principle, we can imagine preparing a state $\mathcal{s}_0$, starting a process $\mathcal{p}_1$ and, while in transition, switch to process $\mathcal{p}_2$.
 At that point, the final state may have characteristics that are different from having run the two process sequentially and that would be evidence of new physics.
-This may correspond to having an atom absorb a photon while it was already in the process of emitting one.
-Or it may correspond to measure the horizontal spin of a particle while we where in the process of measuring the vertical one.
-Whether anything along these lines is actually practicable is unclear, but it's one category.
+Hypothetically, this may correspond to having an atom absorb a photon while it was already in the process of emitting one.
+Or it may correspond to polarizing a particle horizontally while it was in the process of polarizing vertically.
+Whether anything along these lines is actually practicable is unclear, but it's one general strategy.
 
-Another venue would be to find a process that under some conditions is well described by unitary evolution while in others by a projection.
+%Another one would be to find a process that under some conditions is well described by unitary evolution while in others by a projection.
 We may find that at the boundary of the transition the process will be described by neither.
 This may correspond, for example, to those cases where an energy threshold has to be passed to see particle productions or particle absorption.
-It is not clear what kind of precision would be required to elicit a different response, but it's another category.
+It is not clear what kind of precision would be required to elicit a different response, but it's general strategy.
 
-The second main consequence is that to get a more fundamental theory we have to change the state space.
+%The second main conclusion is that to get a more fundamental theory we have to change the state space.
 As quantum states only characterize the equilibria for the projections, we need a different state space for the out-of-equilibrium states.
@@
 We may have to substitute complex Hilbert spaces with something else.
-Given that we have no consensus on why those spaces are needed in the first place, and on whether all the elements are actually physical, it is unclear what we would substitute them with.
+Given that we don't even have consensus on why those spaces are needed in the first place, and on whether all the objects in those spaces are actually physical, it is unclear what we would substitute them with.
 
-But it is actually worse than that.
-If we want a theory that works at the smallest time resolution, it is obvious that our $\mathcal{P}$ would correspond to those processes.
+%We may think we can use our intuition to guide us but it is more likely that it will lead us astray.
+Our ideas of conserved charges, symmetries, quantum numbers and the like may not apply in those out-of-equilibrium states.
+As we know that, during those processes, particles can transform from one type to another, it would make sense to assume they are all made of the same substance and that particles are more or less stable configurations of that substance.
+If that is the case, mass, charge and other quantum numbers may simply be the properties of those stable configurations that have no meaning in between, much like pressure or temperature for an out-of-equilibrium volume of gas.
+Even the symmetries would be qualifying the homogeneity of the equilibrium states.
+
+%Unfortunately, it is probably worse than that.
+Suppose we want a theory that works at the smallest time resolution, at Planck scale.
+Then our set of processes $\mathcal{P}$ would correspond to those processes bounded by that time scale.
 Those will correspond to the basic building blocks with no further refinement possible.
-But in those case the temporal scale of the time evolution is the same scale of the processes: we cannot see time evolution $\mathcal{E}$ as a continuous sequence of processes $\mathcal{p}_t$.
-In fact, there is nothing that would prevent those processes to even be sequential.
+If we now want the law of evolution $\mathcal{E}$ as well, the time scale of $\mathcal{E}$ will be the time scale of the processes $\mathcal{P}$: we cannot see time evolution $\mathcal{E}$ as a continuous sequence of processes $\mathcal{p}_t$.
+As we can no longer pretend that between two fast time scale processes there is always another one, we now must face the fact that these may be running in different regions that may or may not overlap in space and time.
+That is, the execution of there processes is not strictly sequential and therefore not linearly ordered, in which case time cannot be modeled by a number.
 
-In other words, it is possible that we may be very very far from having the correct mathematical tools to be able to characterize those regimes. All
+%If Hilbert spaces are not suitable to characterize states, or if real and rational numbers are not suitable to characterize time, it would mean that we may be very very far from even having the correct mathematical tools to be able to characterize those regimes.
 

### 2019-04-01 Articles/2019-ForQuantumOptics/article.tex
Adding Christine's comments on intro and conclusion

@@
 Assuming a quasistatic evolution, then, means the equilibria of the faster scale processes will evolve slowly, giving rise to continuous and unitary evolution.
-These ideas are enough to get the general form of the projection postulate and the time evolution of quantum mechanics.
+These ideas are sufficient to get the general form of the projection postulate and the time evolution of quantum mechanics.
 Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
@@
 
-In an effort to better understand why the frameworks of classical and quantum mechanics are what they are, we have worked on identifying a minimal set of physical assumptions from which both can be derived.
-This gave us a new perspective that is somewhat unusual among people who fork on fundamental issues in physics.
-The typical view is that the theories are so different that they require a radical different set of starting point, and that classical mechanics in particular should only be viewed as an approximation of the other.
-In this view, the similarities between the theories are due to one being an approximation of the other.
+In an effort to better understand why the frameworks of classical and quantum mechanics are what they are, we have worked on identifying a minimal set of physical assumptions from which both can be derived.[cite] This gave us a new perspective that is somewhat unusual among people who work on fundamental questions in physics.
+The typical view is that the theories are so different that they require a radically different set of starting points, and that classical mechanics in particular should only be viewed as an approximation of quantum mechanics.
+In this view, the similarities between the theories stem from one being an approximation of the other.
 
-We believe a better view is to think of both classical and quantum mechanics as a specialization of a simpler more fundamental theory.
-This would deal with states and laws of evolution as abstract objects, formalizing what are the \emph{necessary} conditions for them to be operationally defined such that they can be experimentally studied.
+We believe a more encompassing view is to consider both classical and quantum mechanics as specializations of a simpler, more fundamental theory.
+This more fundamental theory would deal with states and laws of evolution as abstract objects, formalizing the \emph{necessary} conditions for them to be operationally defined such that they can be experimentally studied.
 The other physical theories (classical and quantum mechanics, thermodynamics and so on) would specialize those concepts to better define what states we are studying and under what conditions.
 In this view, the similarities between the theories are due to the general concepts of the more fundamental theory.
-This different approach can help us pinpoint what elements are specific to what theory, clarify their realm of applicability and, hopefully, give us ideas where new physics may lie.
+This alternative approach can help us pinpoint the elements specific to each theory, clarify their realm of applicability and, hopefully, provide insights into where new physics may lie.
 
-In this paper we start clarifying what are the fundamental relationships and, in particular, how they work in the context of quantum mechanics.
-The general idea is that for a state to be experimentally prepared, manipulated and measured, it needs to be ``stable'' for a maybe very short but certainly finite amount of time.
+In this paper we outline the fundamental elements of a generic model for a physical system, their relationships and, in particular, how they work in the context of quantum mechanics.
+The general idea is that for a state to be experimentally prepared, manipulated and measured, it needs to be ``stable'' for what may be a very short but certainly finite amount of time.
 That is, it must be a, possibly metastable, equilibrium of faster scale processes.
 But a process for which all outputs are equilibria is idempotent (i.e.
-gives the same output no matter how many is applied) and is therefore mathematically modeled by a projection operation on the space of statistical distributions.
+gives the same output no matter how many times it is applied) and is therefore mathematically modeled by a projection operation on the space of statistical distributions.
 We identify quantum projections as these processes.
-If we want a deterministic law, then, it has to be a quasi-static evolution in terms of those fast scale processes.
-As the change has to be small, this leads to continuous evolution over the state space and unitary evolution in the space of statistical distributions.
-We identify unitary evolution as these processes.
+If we want a deterministic law, then, it must be a quasistatic evolution in terms of those faster scale processes.
+As the change must be small, this leads to continuous evolution over the state space and unitary evolution in the space of statistical distributions.
+We identify unitary evolution as this deterministic law.
 
-The necessity of states to be stable for a certain minimum timescale applies to all physical theories, and a deterministic law, then, can only be valid \emph{above} that timescale within each theory.
-In other words, fast scale processes define the state space as their equilibria and the minimum scale $dt$ for the law of evolution.
+The necessity of states to be stable for a certain minimum timescale applies to all physical theories, and a deterministic law, then, can only be valid at \emph{coarser} time resolution within each theory.
+In other words, faster scale processes define the state space as their equilibria and the minimum scale $dt$ for the law of evolution.
 It will be instructive, then, to see how these concepts apply to classical mechanics and thermodynamics as well.
-This will foster the idea that looking at the foundations of all physical theories, and not on quantum mechanics in isolation, fosters a better general understanding of fundamental concepts.
+This will illustrate how looking at the foundations of all physical theories, and not at quantum mechanics in isolation, fosters a better general understanding of fundamental concepts.
 
 In section 2 we will first take the basic framework of quantum mechanics and dress it in a slightly different language.
-The purpose is to make the interplay between pure states, statistical distributions of pure states, projections and unitary evolution more clear.
-In that context it will be easier to see how unitary evolution can be seen as a quasi-static process of infinitesimally close projections.
-In section 3 we will see how the same mathematical structure can be justified from more general physical consideration and how it necessarily applies to all physical theories, including classical mechanics in the thermodynamics.
+The purpose is to make the interplay between pure states, statistical distributions of pure states, projections and unitary evolution clearer.
+In that context it will be easier to see how unitary evolution can be considered a quasistatic process of infinitesimally close projections.
+In section 3 we will see how the same mathematical structure can be justified from more general physical considerations and how it necessarily applies to all physical theories, including classical mechanics and thermodynamics.
 
@@
 We will then recover unitary evolution (i.e.
-the Schroedinger equation) as the quasi-static deterministic evolution given by continuous projections.
+the Schroedinger equation) as the quasistatic deterministic evolution given by continuous projections.
 To do so, we will introduce a different mathematical notation which will make explicit the difference between pure states, statistical distributions and mixed states, to help dispel potential confusion.
@@
 
-If we accept this premise, then the natural thing to do is to write the conditions for a quasi-static deterministic evolution.
+If we accept this premise, then the natural thing to do is to write the conditions for a quasistatic deterministic evolution.
 That is, we assume that at each moment the system is in equilibrium and we change the state just a little so that each pure state is mapped to a single well-defined pure state.
@@
 This requires the evolution to be unitary and we can use standard techniques from here, which will lead to the Schroedinger equation.
-In other words, \textbf{the Schroedinger equation is the quasi-static deterministic evolution of a process of continuous projections}.
-This approach, therefore, proceeds in the opposite way of others: instead of assuming the Schroedinger equation is always valid and deriving projections as a special case, we assume projections are valid approximations of faster scale processes and the Schroedinger equation is the special quasi-static deterministic case.
+In other words, \textbf{the Schroedinger equation is the quasistatic deterministic evolution of a process of continuous projections}.
+This approach, therefore, proceeds in the opposite way of others: instead of assuming the Schroedinger equation is always valid and deriving projections as a special case, we assume projections are valid approximations of faster scale processes and the Schroedinger equation is the special quasistatic deterministic case.
 
@@
 We also note that in quantum field theory, to calculate cross-sections and particle decays, one puts the initial states at minus infinity and the final states at plus infinity for processes that, in reality, take a tiny fraction of a second.
-This is reminiscent of a quasi-static approximation.
-Finally, while the approach presented here does not solve the so called ``measurement problem'', it does however give us an understanding of why we have the problem to begin with: our states are really just the equilibrium states of faster scale processes and those cannot, in general, be described by quasi-static ones.
+This is reminiscent of a quasistatic approximation.
+Finally, while the approach presented here does not solve the so called ``measurement problem'', it does however give us an understanding of why we have the problem to begin with: our states are really just the equilibrium states of faster scale processes and those cannot, in general, be described by quasistatic ones.
 
@@
 Yet, the overall structure remains the same: we have a timescale $dt$ such that within each interval $[t-dt, t]$ there is a process $\mathcal{p}_t$ such that the state $\mathcal{s}_t$ is an equilibrium of $\mathcal{p}_t$.
-The evolution $\mathcal{E}$, then, is quasi-static in terms of the processes in $\mathcal{P}$.
-That is, \textbf{all deterministic laws are quasi-static evolutions in terms of system-environment interaction processes}.
+The evolution $\mathcal{E}$, then, is quasistatic in terms of the processes in $\mathcal{P}$.
+That is, \textbf{all deterministic laws are quasistatic evolutions in terms of system-environment interaction processes}.
 
@@
 Like in thermodynamics, we have states that are equilibria of non-deterministic processes, of which we are not able to give a more fundamental description.
-Quantities are only well defined at equilibrium and the law of evolution is the quasi-static evolution where we assume the system changes slowly and we are always at equilibrium.
+Quantities are only well defined at equilibrium and the law of evolution is the quasistatic evolution where we assume the system changes slowly and we are always at equilibrium.
 What is interesting is that \textbf{in quantum mechanics what drives the deterministic evolution are the same processes that drive the equilibrium}.
@@
 
-We have seen that quantum mechanics fits in a more general framework that applies to all other physical theories.
-States are always equilibria of some faster scale process, which in quantum mechanics are modeled with the projection.
-States are chosen to satisfy a law of evolution, which is necessarily a quasi-static process in terms of the faster scale processes, which in quantum mechanics is modeled by the Schroedinger equation.
+We have seen that quantum mechanics fits within a more general framework that applies to all physical theories.
+States are always equilibria of faster scale processes, which in quantum mechanics are modeled with the projection.
+States are chosen to satisfy a deterministic law of evolution, necessarily describing a quasistatic process, which in quantum mechanics is modeled by the Schroedinger equation.
 This general framework only uses standard physics concepts and puts no special role on the observer except the choice of what processes the system will be subjected to.
 
-Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasi-static.
-If we were to implement a measurement process, then, in a way that the transfer of information between the system and the probe happens in a slow and controlled fashion, this would will not provide a violation to the Schroedinger equation, nor it would tell us that all projections are quasi-static processes.
-A better description of the faster scale processes given by the projections, then, will inevitable need a new state space as the standard quantum states only describes equilibria, and we should expect the faster scale processes to go out of equilibria.
+Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
+If we were to implement a measurement process, then, in a way that the transfer of information between the system and the probe happens in a slow and controlled fashion, this could not provide a violation of the Schroedinger equation, nor would it tell us that all projections are quasistatic processes.
+A better description of the faster scale processes given by the projections, then, will inevitably need a new state space as the standard quantum states only describe equilibria, and we should expect the faster scale processes to go out of equilibrium.
+Ultimately, a description at Plank time scale will fail to be continuous and quasistatic since a finite change in a finite time will not be divided into infinitesimal adiabatic changes over infinitely many time steps.
 
-To understand what is the right mathematical framework for those regimes, we are proceeding by first identifying the physical requirements for all mathematical structures already used in physics, starting from the most basic ones like topological spaces (upon which is based all of differential geometry) and $\sigma$-algebras (upon which is based all of measure theory and probability theory).
+To understand the suitable mathematical framework for those regimes where unitary evolution may not hold, we are proceeding by first identifying the physical requirements for all mathematical structures already used in physics, starting from the most basic ones like topological spaces (upon which is based all of differential geometry) and $\sigma$-algebras (upon which is based all of measure theory and probability theory).
 This long term effort is centered mainly around the development of an open access book in which these issues can be explored in full detail and with a rigor more typical of mathematical disciplines.
-This rigor is what forces us to spell out the often hidden assumptions we have while constructing physical theories.
-By making explicit under which physically motivated circumstances time can be modeled by real numbers, states can be modeled by classical phase space or by quantum Hilbert spaces, we believe we will be in a better position to understand what happens when we can't.
+This rigor forces us to spell out the often hidden assumptions we have while constructing physical theories.
+By making explicit, for example, under which physically motivated circumstances time can be modeled by real numbers, or when states can be represented within Hilbert spaces, we believe we will be in a better position to understand what happens when those assumptions fail.
 
@@
 %The first main conclusion is that we should be able to find processes that are neither described by unitary evolution nor by the projection.
-These would correspond to processes for which the out-of-equilibrium dynamics becomes relevant so the quasi-static evolution is not a good approximation and neither is the instantaneous transition.
+These would correspond to processes for which the out-of-equilibrium dynamics becomes relevant so the quasistatic evolution is not a good approximation and neither is the instantaneous transition.
 Even if we do not have a theory for those regimes, we can, at least conceptually, discuss a couple of general strategies.
@@
 %For example, we can note that slow changing conditions are unlikely to reveal new dynamics.
-As long as the system evolves continuously, the quasi-static approximation will hold.
+As long as the system evolves continuously, the quasistatic approximation will hold.
 It is also conceivable that, in some conditions, a process that may look like a projection end-to-end is actually well described by a slow transition.
-In those cases the quasi-static approximation evidently holds and no new phenomena would be observed.
+In those cases the quasistatic approximation evidently holds and no new phenomena would be observed.
 What is the right time scale, though, is not clear to us since ``slow'' in this setting is still ``very fast''.

### 2019-04-02 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
Aligned definitions with book

@@
 
+\def\Bool{\mathbb{B}}
 \def\TRUE{\textsc{true}}
@@
 \newcommand{\obs}[1][s] {\mathsf{#1}}
+
 % Formatting for possibilities
+\def\logCtx{\mathcal{S}}
+\newcommand{\pAss}[1][\mathcal{S}] {\mathcal{A}_{#1}}
 
@@
 	A \textbf{statement} $\stmt$ is an assertion that is either true or false.
-Formally, a statement is an element of the set $\mathcal{S}$ of all statements upon which is defined a function $\truth: \mathcal{S} \to \mathbb{B}$ that returns the truth value for each element.
+A \textbf{logical context} $\logCtx$ is a collection of statements with well defined logical relationships.
+Formally, a logical context $\logCtx$ is a collection of elements called statements upon which is defined a function $\truth: \logCtx \to \Bool$.
 \end{axiom}
@@
 But our scientific model may constrain certain statements or statement combinations to be ruled out.
-For example, the statements \statement{this animal is a dog} and \statement{this animal is a cat} can both be either true or false, but the statement \statement{this animal is a cat and a dog} can never be true.
+For example, the statements \statement{this animal is a dog} and \statement{this animal is a cat} can both be either true or false, but they can't both be true.
 That is, the role of our logic system is not to keep track of what is true and false, but of what cannot possibly be true and cannot possibly be false.
@@
 
-\begin{defn}
-	Given a collection of statements $\{\stmt_i\}^n_{i=1}$, a \textbf{consistent truth assignment} is a collection of truth values $\{t_i\}^n_{i=1}$ such that it is logically consistent to simultaneously suppose that $\truth(\stmt_i) = t_i$ for all $1 \leq i \leq n$.
-That is, from those assumptions it cannot be proven that $\truth(\stmt_i) \neq t_i$ for any $1 \leq i \leq n$.
-This definition generalizes to the case of infinite, possibly uncountable, indexed families.
-\end{defn}
+\begin{axiom}\label{ax_possible_assignemetns}
+	A \textbf{possible assignment} for a logical context $\logCtx$ is a map $a : \logCtx \to \Bool$ that assigns a truth value to each statement in a way consistent with the content of the statements.
+Formally, each logical context comes equipped with a set $\pAss \subseteq \Bool^{\logCtx}$ such that $\truth \in \pAss$.
+A map $a : \logCtx \to \Bool$ is a possible assignment for $\logCtx$ if $a \in \pAss$.
+\end{axiom}
 
-Note that the truth assignments are just hypotheticals.
+Note that the possible assignments are just hypotheticals.
 They are not actual physical entities.
@@
 Logical consistency is defined on the truth assignments and not on the truth values.
-Therefore we need a way to track whether the meaning of each statement allows it to be true or not.
-
-\begin{axiom}\label{ax_possibilities}
-	The \textbf{possibilities} of a statement $\stmt$ are the possible truth values allowed by the content of the statement.
-Formally, on $\mathcal{S}$ is also defined a function $\possFn: \mathcal{S} \to \{\{\FALSE, \TRUE\},\{\FALSE\},\{\TRUE\}\}$ such that:
-	\begin{itemize}
-		\item $\truth(\stmt) \in \possFn(\stmt)$ for all $\stmt \in \mathcal{S}$. This remains valid in every consistent truth assignment.
-		\item for any collection of statements $\{\stmt_i\}^n_{i=1}$, for any $1 \leq j \leq n$ and for any $t \in \possFn(\stmt_j)$ there exists a consistent truth assignment $\{t_i\}^n_{i=1}$ such that $t_j = t$.
-This generalizes to the case of infinite, possibly uncountable, indexed families.
-	\end{itemize}
-\end{axiom}
-
-With this axiom, we can distinguish between statements that can never be true and those that just happen to be true.
+This allows us, for example, to track whether the meaning of each statement allows it to be true or not.
 
 \begin{defn}
-	A \textbf{tautology} $\tautology$ is a statement that must be true simply because of its content. That is, $\possFn(\tautology) = \{\TRUE\}$.
+	A \textbf{tautology} $\tautology$ is a statement that is true in every possible assignment.
 \end{defn}
@@
 \begin{defn}
-	A \textbf{contradiction} $\contradiction$ is a statement that must be false simply because of its content. That is, $\possFn(\contradiction) = \{\FALSE\}$.
+	A \textbf{contradiction} $\contradiction$ is a statement that is false in every possible assignment.
 \end{defn}
@@
 	We can always construct a statement whose truth value arbitrarily depends on an arbitrary set of statements.
-Formally, given an arbitrary truth function $f_{\mathbb{B}} : \mathbb{B}^n \to \mathbb{B}$ there exists a function $f : \mathcal{S}^n \to \mathcal{S}$ such that
-	$$\truth(f(\stmt_1, ..., \stmt_n)) = f_{\mathbb{B}}(\truth(\stmt_1), ..., \truth(\stmt_n))$$
-	and the same relationship remains valid in every consistent truth assignment. This also holds in the case of infinite, possibly uncountable, arguments.
+Formally, let $S \subseteq \logCtx$ be a set of statements and $f_{\Bool} : \Bool^S \to \Bool$ an arbitrary function from an assignment of $S$ to a truth value.
+Then we can always find a statement $\bar{\stmt} \in \logCtx$ such that
+	$$a(\bar{\stmt}) = f_{\Bool}(\{a(\stmt)\}_{\stmt \in S})$$
+	for every possible assignment $a \in \pAss$.
 \end{axiom}
@@
 	Two statements $\stmt_1$ and $\stmt_2$ are \textbf{equivalent} $\stmt_1 \equiv \stmt_2$ if they must be equally true or false simply because of their content.
-Formally, $\stmt_1 \equiv \stmt_2$ if and only if $(\stmt_1 \AND \stmt_2) \OR (\NOT\stmt_1 \AND \NOT\stmt_2)$ is a tautology.
+Formally, $\stmt_1 \equiv \stmt_2$ if and only if $a(\stmt_1) = a(\stmt_2)$ for all $a \in \pAss$.
 \end{defn}
 
-From this definition, one can prove that two statements are equivalent if and only if they have the same truth in all consistent truth assignments.
-From that one can prove the following two propositions.
+From this definition one can prove the following two propositions.
 
@@
 
-This system not only allows us to use formal classical logic while keeping the statements informal, but is also equipped to capture causal relationships.\footnote{Note that the axioms prevent any form of paradox, simply because the possibilities of a statement are never empty and a consistent truth assignment must exist.
-Therefore situations like \statement{this statement is not true} are ruled out simply because they cannot satisfy the axioms.} For example, $\stmt_1=$\statement{the thermometer indicator is between 24 C and 25 C} and $\stmt_2=$\statement{the temperature is between 24 C and 25 C} are equivalent because both $\stmt_1 \AND \NOT \stmt_2$ and $\NOT \stmt_1 \AND \stmt_2$ are contradictions (on the assumption that our thermometer is actually working).
+This system not only allows us to use formal classical logic while keeping the statements informal, but is also equipped to capture causal relationships.\footnote{Note that the axioms prevent any form of paradox, simply because a possible assignment must exist.
+Therefore situations like \statement{this statement is not true} are ruled out simply because they cannot satisfy the axioms.} For example, $\stmt_1=$\statement{the thermometer indicator is between 24 C and 25 C} and $\stmt_2=$\statement{the temperature is between 24 C and 25 C} are equivalent because one is true if and only if the other is true (on the assumption that our thermometer is actually working).
 We can also define other semantic relationships.
@@
 	\begin{itemize}
-		\item $\stmt_1$ \textbf{is narrower than} $\stmt_2$ (noted $\stmt_1 \narrower \stmt_2$) if $\stmt_2$ is true whenever $\stmt_1$ is true simply because of their content.
-That is, $\stmt_1 \AND \NOT \stmt_2 \equiv \contradiction$.
+		\item $\stmt_1$ \textbf{is narrower than} $\stmt_2$ (noted $\stmt_1 \narrower \stmt_2$) if $\stmt_2$ is true whenever $\stmt_1$ is true.
 		\item $\stmt_1$ \textbf{is broader than} $\stmt_2$ (noted $\stmt_1 \broader \stmt_2$) if $\stmt_2 \narrower \stmt_1$.
-		\item $\stmt_1$ \textbf{is compatible to} $\stmt_2$ (noted $\stmt_1 \comp \stmt_2$) if their content allows them to be true at the same time. That is, $\stmt_1 \AND \stmt_2 \nequiv \contradiction$.
+		\item $\stmt_1$ \textbf{is compatible to} $\stmt_2$ (noted $\stmt_1 \comp \stmt_2$) if there is an assignment in which they are both true.
 
@@
 \begin{defn}
-	The elements of a set of statements $S \subseteq \mathcal{S}$ are said to be \textbf{independent} (noted $\stmt_1 \indep \stmt_2$ for a set of two) if their content is such that any combination of their possibilities is allowed.
-That is, $\possFn(f(S)) = f(\bigtimes\limits_{\stmt \in S} \possFn(\stmt))$ for any truth function $f : \mathbb{B}^{|S|} \to \mathbb{B}$.
-The negation of independence will be noted by $\nindep$.
+	The elements of a set of statements $S \subseteq \logCtx$ are said to be \textbf{independent} (noted $\stmt_1 \indep \stmt_2$ for a set of two) if the assignment of any subset of statements does not depend on the assignment of the others.
+That is, a set of statements $S \subseteq \logCtx$ is independent if given a family  $\{t_{\stmt}\}_{\stmt \in S}$ such that each $t_{\stmt} \in \Bool$ is a possible assignment for the respective $\stmt$ we can always find $a \in \pAss$ such that $a(\stmt) = t_{\stmt}$ for all $\stmt \in S$.
 \end{defn}
@@
 In principle, given a set of verifiable statements we can simply start testing them one after the other.
-However, if we are given a set of uncountable statements then, even if we have an indefinitely long time at our disposal, we will not be able to create a procedure that eventually tests all statements.
+However, if we are given a set of uncountable statements then, even if we have an indefinitely long time at our disposal, we will not be able to create a procedure that eventually reaches all statements.
 However, we may not need to actually run all tests for all statements.

### 2019-04-02 Articles/2019-ForQuantumOptics/article.tex
Minor change

@@
 A better description of the faster scale processes given by the projections, then, will inevitably need a new state space as the standard quantum states only describe equilibria, and we should expect the faster scale processes to go out of equilibrium.
-Ultimately, a description at Plank time scale will fail to be continuous and quasistatic since a finite change in a finite time will not be divided into infinitesimal adiabatic changes over infinitely many time steps.
+Consequently, a description at Planck time scale will ultimately fail to be continuous and quasistatic since a finite change in a finite time will not be divided into infinitesimal adiabatic changes over infinitely many time steps.
 

### 2019-04-19 Articles/2019-ForQuantumOptics/MDPI_template/template.tex
Finalized version for submission

--- 2019-04-01 Articles/2019-ForQuantumOptics/article.tex
@@
-\documentclass[11pt]{article}
+%  LaTeX support: latex@mdpi.com 
+%  In case you need support, please attach all files that are necessary for compiling as well as the log file, and specify the details of your LaTeX setup (which operating system and LaTeX version / tools you are using).
 
-\usepackage[margin=1.25in]{geometry}
+%=================================================================
+\documentclass[applsci,article,submit,moreauthors,pdftex]{Definitions/mdpi}
 
+% If you would like to post an early version of this manuscript as a preprint, you may use preprint as the journal and change 'submit' to 'accept'.
+The document class line would be, e.g., \documentclass[preprints,article,accept,moreauthors,pdftex]{mdpi}.
+This is especially recommended for submission to arXiv, where line numbers should be removed before posting.
+For preprints.org, the editorial staff will make this change immediately prior to posting.
+
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% acoustics, actuators, addictions, admsci, aerospace, agriculture, agriengineering, agronomy, algorithms, animals, antibiotics, antibodies, antioxidants, applsci, arts, asc, asi, atmosphere, atoms, axioms, batteries, bdcc, behavsci , beverages, bioengineering, biology, biomedicines, biomimetics, biomolecules, biosensors, brainsci , buildings, cancers, carbon , catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, children, cleantechnol, climate, clockssleep, cmd, coatings, colloids, computation, computers, condensedmatter, cosmetics, cryptography, crystals, dairy, data, dentistry, designs , diagnostics, diseases, diversity, drones, econometrics, economies, education, electrochem, electronics, energies, entropy, environments, epigenomes, est, fermentation, fibers, fire, fishes, fluids, foods, forecasting, forests, fractalfract, futureinternet, futurephys, galaxies, games, gastrointestdisord, gels, genealogy, genes, geohazards, geosciences, geriatrics, hazardousmatters, healthcare, heritage, highthroughput, horticulturae, humanities, hydrology, ijerph, ijfs, ijgi, ijms, ijns, ijtpp, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jcdd, jcm, jcp, jcs, jdb, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmse, jnt, jof, joitmc, jpm, jrfm, jsan, land, languages, laws, life, literature, logistics, lubricants, machines, magnetochemistry, make, marinedrugs, materials, mathematics, mca, medicina, medicines, medsci, membranes, metabolites, metals, microarrays, micromachines, microorganisms, minerals, modelling, molbank, molecules, mps, mti, nanomaterials, ncrna, neuroglia, nitrogen, notspecified, nutrients, ohbm, particles, pathogens, pharmaceuticals, pharmaceutics, pharmacy, philosophies, photonics, physics, plants, plasma, polymers, polysaccharides, preprints , proceedings, processes, proteomes, psych, publications, quantumrep, quaternary, qubs, reactions, recycling, religions, remotesensing, reports, resources, risks, robotics, safety, sci, scipharm, sensors, separations, sexes, signals, sinusitis, smartcities, sna, societies, socsci, soilsystems, sports, standards, stats, surfaces, surgeries, sustainability, symmetry, systems, technologies, test, toxics, toxins, tropicalmed, universe, urbansci, vaccines, vehicles, vetsci, vibration, viruses, vision, water, wem, wevj
+
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, benchmark, book, bookreview, briefreport, casereport, changes, comment, commentary, communication, conceptpaper, conferenceproceedings, correction, conferencereport, expressionofconcern, extendedabstract, meetingreport, creative, datadescriptor, discussion, editorial, essay, erratum, hypothesis, interestingimages, letter, meetingreport, newbookreceived, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, supfile, technicalnote, viewpoint
+% supfile = supplementary materials
+
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. If eps figures are used, remove the option pdftex and use LaTeX and dvi2pdf.
+
+\usepackage{braket}
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage{graphicx}
 \usepackage{amsfonts}
 \usepackage{dutchcal}
-\usepackage{braket}
-\usepackage{enumitem}
 
-\usepackage{tikz}
-\usepackage{forest}
-\usetikzlibrary{trees}
-\usetikzlibrary{calc}
-\usepackage{calculator}
-\usepackage{standalone}
+%=================================================================
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother
+\pubvolume{xx}
+\issuenum{1}
+\articlenumber{5}
+\pubyear{2019}
+\copyrightyear{2019}
+%\externaleditor{Academic Editor: name}
+\history{Received: date; Accepted: date; Published: date}
+%\updates{yes} % If there is an update available, un-comment this line
 
-\usepackage{braket}
+%% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
+%\continuouspages{yes}
 
-\begin{document}
+%------------------------------------------------------------------
+% The following line should be uncommented if the LaTeX file is uploaded to arXiv.org
+%\pdfoutput=1
 
-\title{Unitary evolution as a quasistatic process of infinitesimal collapses}
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, calc, indentfirst, fancyhdr, graphicx, lastpage, ifthen, lineno, float, amsmath, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, amsthm, hyphenat, natbib, hyperref, footmisc, geometry, caption, url, mdframed, tabto, soul, multirow, microtype, tikz
 
-\maketitle
+%=================================================================
+%% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
 
-\begin{abstract}
-Any state, to be properly defined, has to be an equilibrium of all processes happening at a faster timescale.
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Unitary evolution as a quasistatic process of infinitesimal collapses}
+
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0002-1071-6251} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0001-9540-4988} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Gabriele Carcassi $^{1}$\orcidA{}* and Christine A. Aidala $^{1}$\orcidB{}}
+
+% Authors, for metadata in PDF
+\AuthorNames{Gabriele Carcassi and Christine Aidala}
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+$^{1}$ \quad Department of Physics, University of Michigan, Ann Arbor, MI 48109, USA}
+
+% Contact information of the corresponding author
+\corres{Correspondence: carcassi@umich.edu}
+
+% Current address and/or shared authorship
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{Any state, to be properly defined, has to be an equilibrium of all processes happening at a faster timescale.
 As processes with equilibria are idempotent, they can be expressed as projections in a probability space.
 Assuming a quasistatic evolution, then, means the equilibria of the faster scale processes will evolve slowly, giving rise to continuous and unitary evolution.
-These ideas are sufficient to get the general form of the projection postulate and the time evolution of quantum mechanics.
-Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
-\end{abstract}
+In this paper we show that these ideas are sufficient to get the general form of the projection postulate and the time evolution of quantum mechanics, while also creating a model that applies to other physical theories, such as classical mechanics and thermodynamics.
+In this light, the different theories are specializations of a broader model.
+As unitary evolution corresponds to quasistatic evolution, violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
+}
 
-%\tableofcontents
-%\newpage
+% Keywords
+\keyword{foundations of quantum mechanics; quasistatic evolution; processes and equilibria}
 
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Diversity
+%\LSID{\url{http://}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences:
+%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data:
+%\dataset{DOI number or link to the deposited data set in cases where the data set is published or set to be published separately.
+If the data set is submitted and will be published as a supplement to this paper in the journal Data, this field will be filled by the editors of the journal.
+In this case, please make sure to submit the data set as a supplement when entering your manuscript into our manuscript editorial system.}
+
+%\datasetlicense{license under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Toxins
+%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
+
+%\setcounter{secnumdepth}{4}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
 
-In an effort to better understand why the frameworks of classical and quantum mechanics are what they are, we have worked on identifying a minimal set of physical assumptions from which both can be derived.[cite] This gave us a new perspective that is somewhat unusual among people who work on fundamental questions in physics.
+In an effort to better understand why the frameworks of classical and quantum mechanics are what they are, we have worked on identifying a minimal set of physical assumptions from which both can be derived.\cite{Carc1} This gave us a new perspective that is somewhat unusual among people who work on fundamental questions in physics.
 The typical view is that the theories are so different that they require a radically different set of starting points, and that classical mechanics in particular should only be viewed as an approximation of quantum mechanics.
@@
 
-In this paper we outline the fundamental elements of a generic model for a physical system, their relationships and, in particular, how they work in the context of quantum mechanics.
-The general idea is that for a state to be experimentally prepared, manipulated and measured, it needs to be ``stable'' for what may be a very short but certainly finite amount of time.
+In this paper we outline the fundamental elements of a general model for a physical system, their relationships and, in particular, how they work in the context of quantum mechanics.
+The basic idea is that for a state to be experimentally prepared, manipulated and measured, it needs to be ``stable'' for what may be a very short but certainly finite amount of time.
 That is, it must be a, possibly metastable, equilibrium of faster scale processes.
@@
 Therefore we define the set of mixed states $\mathcal{M}$ as the set of statistical ensembles that are physically different, which is usually represented by a density matrix.
-Mathematically, we can also represent physical equivalence between statistical ensembles with the equivalence operator $\mathcal{d}_1 \equiv \mathcal{d}_2$ and the set of mixed states $\mathcal{M} = \mathcal{D}_{/_\equiv}$ will be the quotient space.
+Mathematically, we can also represent physical equivalence between statistical ensembles with an equivalence operator $\mathcal{d}_1 \equiv \mathcal{d}_2$ and the set of mixed states $\mathcal{M} = \mathcal{D}_{/_\equiv}$ will be the quotient space.
 
@@
 This is reminiscent of a quasistatic approximation.
-Finally, while the approach presented here does not solve the so called ``measurement problem'', it does however give us an understanding of why we have the problem to begin with: our states are really just the equilibrium states of faster scale processes and those cannot, in general, be described by quasistatic ones.
+
+Finally, we note that the approach presented here does not solve the so called ``measurement problem''\cite{Genovese,Bassi}, at least not in the way it is typically framed.
+We still do not have a description of what happens during an arbitrary process (e.g.
+a projection), yet we have an explanation of why we cannot provide one in terms of a quasistatic evolution (i.e.
+unitary evolution) since a special case cannot describe the general one.\footnote{We purposely leave unspecified whether the faster scale processes are driven by internal or external dynamics, or by a mixture of the two.
+As such, this approach does not technically pick an interpretation: it merely states that whatever happens during a measurement in a given interpretation is also happening at every instant in time.}
+
+%since the fast out-of-equilibrium process cannot be described with a quasi-static evolution.
 
@@
 Under those conditions, no interactions with the system are possible.
-Photons cannot scatter off of it, nor gravitational waves; no way to synchronize clocks between the system and the environment, no way to tell where the system is with respect to the environment.
+Photons cannot interact with it, nor gravitational waves; no way to synchronize clocks between the system and the environment, no way to tell where the system is with respect to the environment.
 To us in the environment, within those circumstances, the system has vanished.
@@
 This means that we can never have a complete independent description of the system.\footnote{Ultimately, this is why classical mechanics conceptually fails.
-See XXX for more discussion.} We need to leave out some ``unstated part'' whose future will depend on the environment.
+See \cite{Carc1} for more discussion.} We need to leave out some ``unstated part'' whose future will depend on the environment.
 In other words, \textbf{the system always has a boundary with the environment and, through it, is in constant interaction}.
 
-This unavoidable interaction with the environment also comes into play in the definition of a system and its state.
+This unavoidable interaction with the environment comes into play in the definition of a system and its state.
 For example, we can talk about the position of a ball because we sit on the surface of the earth at around 273 K.
@@
 Therefore even if the unstated part is not part of the state, its interaction with the environment has to be ``suitable'' for the state to be defined in the first place.
-In other words, \textbf{a physical property of a system is well defined and measurable if and only if the constant interaction with the environment keeps it that way.}\footnote{A possible misconception here is thinking of a process that does not change anything as ``doing nothing''.}
+In other words, \textbf{a physical property of a system is well defined and measurable if and only if the constant interaction with the environment keeps it that way.}\footnote{A possible misconception is thinking that a process that does not change our state is ``doing nothing''.}
 
@@
 For example, as we change the volume of a gas, the region against the moving wall will be compressed sooner than the opposite region, but if we wait a short while we will again obtain a uniform pressure.
-If we want to study that transition, we may conceptually divide the system into small volumes and the pressure is replaced by a pressure field.
-Still, we need to assume each small volume to have enough molecules to be in equilibrium.
-We can go into more detail, and describe the distribution of position and velocities of the molecules.
+If we want to study that transition, we may conceptually divide the system into small volumes and we replace the pressure by a pressure field.
+Still, we need to assume each small volume has enough molecules to be in equilibrium.
+We can go into more detail, and describe the distribution of position and velocity of the molecules.
 Yet, we need to assume that the molecules themselves are not undergoing chemical changes.
@@
 It is fundamental because it always applies and what changes case by case are the specifics.
-If we study a ball using classical particle mechanics, for example, the position and momentum of the center of mass will be defined in a short time scale integrating the jiggling created by the random collisions of air particles.
+In all cases we are going to have a set of states that are equilibria of faster scale processes.
+Each faster scale process will be, in general, non-deterministic in terms of the state and will act on the space of probability distributions.
+The law of evolution will be a quasistatic process in terms of the faster scale processes.
+
+In classical mechanics, the state of a ball will be given by the position and momentum of the center of mass, which will be defined over a short time scale by integrating the jiggling created by the random collisions of air particles and internal motion of the ball's molecules.
 That averaging process will be the same regardless of the values of position and momentum; all states are equilibria of the same process and therefore $\mathcal{P}$ contains only one element: the identity map.
@@
 
-In this framework all processes in $\mathcal{P}$ always represent abrupt transitions to a new equilibrium.
-In thermodynamics they could represent the crystallization of a supersaturated solution or the ignition of a mixture of oxygen and hydrogen to form water.
-In quantum mechanics they represent the absorption of an incoming photon by a hydrogen atom or the decay of a muon into an electron and two neutrinos.
+In this framework, then, the processes in $\mathcal{P}$ do not represent measurements, they simply represent transitions to a new equilibrium.
+In thermodynamics they may represent the crystallization of a supersaturated solution or the ignition of a mixture of oxygen and hydrogen to form water.
+In quantum mechanics they may represent the absorption of an incoming photon by a hydrogen atom or the decay of a muon into an electron and two neutrinos.
 In all these cases we start from a state of (possibly metastable) equilibrium and we end with a new equilibrium which is not, in general, described in terms of the initial state space.
@@
 We do not believe the similar conceptual structure to be simply a formal analogy but to represent analogue physical mechanisms.
-Quantum projections happen for the same reasons supersaturated solutions crystallize: an abrupt change in external or internal physical conditions.
+Quantum projections happen for the same reasons supersaturated solutions crystallize: a change in external or internal physical conditions that leads to a new equilibrium.
 
@@
 In the case the particle itself is absorbed, or destroyed in any way, then we are already in the realm of fast processes.
-In the other cases, the particle has to initiate a chain reaction that has to transfers energy stored in the detector system to the readout.
+In the other cases, the particle has to initiate a chain reaction that has to transfer energy stored in the detector system to the readout.
 The detector, then, has to be prepared in a state sufficiently stable to not change if no particle interacts with it but unstable enough that a small perturbation makes it transition to a lower energy.
 That is, the detector has to be prepared in a metastable state and the particle interaction has to make it transition to a more stable state.
-But a transition from a metastable state to a more stable state is exactly what our projections are supposed to represent.
-%https://www.researchgate.net/publication/280330121_A_Heuristic_Approach_to_the_Quantum_Measurement_Problem_How_to_Distinguish_Particle_Detectors_from_Ordinary_Objects
+But a transition from a metastable state to a more stable state is exactly what our projections are supposed to represent.\footnote{The role of metastability in detectors has already been discussed, for example, by \cite{Daneri} and \cite{Merlin}.}
+
+On the other hand, not all measurements are projections.
+In a weak measurement\cite{Tamir}, for example, the system interacts with the probe according to unitary evolution (i.e.
+adiabatically) and the probe itself goes through the projection.
+So the association of a measurement with a projection will not always work, while a process with equilibria can always be mathematically formalized as a projection.
 
@@
 
-Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
+Violations of the Schroedinger equation should be expected only for processes that cannot be considered quasistatic.
 If we were to implement a measurement process, then, in a way that the transfer of information between the system and the probe happens in a slow and controlled fashion, this could not provide a violation of the Schroedinger equation, nor would it tell us that all projections are quasistatic processes.
-A better description of the faster scale processes given by the projections, then, will inevitably need a new state space as the standard quantum states only describe equilibria, and we should expect the faster scale processes to go out of equilibrium.
-Ultimately, a description at Plank time scale will fail to be continuous and quasistatic since a finite change in a finite time will not be divided into infinitesimal adiabatic changes over infinitely many time steps.
+On the other hand, a better description of the faster scale processes given by the projections cannot be given in terms of unitary evolution, as this represents the quasistatic case.
+It would inevitably require a new state space as the standard quantum states can only describe equilibria of those processes.
+Moreover, a description at Planck time scale will ultimately fail to be continuous and quasistatic since a finite change in a finite time will not be divisible into infinitesimal adiabatic changes over infinitely many time steps.
 
 To understand the suitable mathematical framework for those regimes where unitary evolution may not hold, we are proceeding by first identifying the physical requirements for all mathematical structures already used in physics, starting from the most basic ones like topological spaces (upon which is based all of differential geometry) and $\sigma$-algebras (upon which is based all of measure theory and probability theory).
-This long term effort is centered mainly around the development of an open access book in which these issues can be explored in full detail and with a rigor more typical of mathematical disciplines.
+This long term effort is centered mainly around the development of an open access book\cite{Carc3} in which these issues can be explored in full detail and with a rigor more typical of mathematical disciplines.
 This rigor forces us to spell out the often hidden assumptions we have while constructing physical theories.
@@
 
-%\section{???}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
 
-%If we accept the line of thinking present in the previous two sections, we can try and draw some conclusions.
-These will be necessary qualitative as much further work would be needed to make any quantitative prediction.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following are available online at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title.}
 
-%The first main conclusion is that we should be able to find processes that are neither described by unitary evolution nor by the projection.
-These would correspond to processes for which the out-of-equilibrium dynamics becomes relevant so the quasistatic evolution is not a good approximation and neither is the instantaneous transition.
-Even if we do not have a theory for those regimes, we can, at least conceptually, discuss a couple of general strategies.
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following are available at \linksupplementary{s1}, Figure S1: title, Table S1: title, Video S1: title. A supporting video article is available at doi: link.}
 
-%For example, we can note that slow changing conditions are unlikely to reveal new dynamics.
-As long as the system evolves continuously, the quasistatic approximation will hold.
-It is also conceivable that, in some conditions, a process that may look like a projection end-to-end is actually well described by a slow transition.
-In those cases the quasistatic approximation evidently holds and no new phenomena would be observed.
-What is the right time scale, though, is not clear to us since ``slow'' in this setting is still ``very fast''.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work reported.}
 
-%The main issue is that we have to catch the system out of equilibrium.
-In principle, we can imagine preparing a state $\mathcal{s}_0$, starting a process $\mathcal{p}_1$ and, while in transition, switch to process $\mathcal{p}_2$.
-At that point, the final state may have characteristics that are different from having run the two process sequentially and that would be evidence of new physics.
-Hypothetically, this may correspond to having an atom absorb a photon while it was already in the process of emitting one.
-Or it may correspond to polarizing a particle horizontally while it was in the process of polarizing vertically.
-Whether anything along these lines is actually practicable is unclear, but it's one general strategy.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\funding{Funding for this work was provided in part by the MCubed program of the University of Michigan.}
 
-%Another one would be to find a process that under some conditions is well described by unitary evolution while in others by a projection.
-We may find that at the boundary of the transition the process will be described by neither.
-This may correspond, for example, to those cases where an energy threshold has to be passed to see particle productions or particle absorption.
-It is not clear what kind of precision would be required to elicit a different response, but it's general strategy.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\acknowledgments{This article is part of a larger project, Assumptions
+	of Physics, that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.}
 
-%The second main conclusion is that to get a more fundamental theory we have to change the state space.
-As quantum states only characterize the equilibria for the projections, we need a different state space for the out-of-equilibrium states.
-This suggests that developing the right physical model for what happens during a projection is not a simple matter of finding a suitable Hamiltonian or a non-linear law of evolution.
-We may have to substitute complex Hilbert spaces with something else.
-Given that we don't even have consensus on why those spaces are needed in the first place, and on whether all the objects in those spaces are actually physical, it is unclear what we would substitute them with.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\conflictsofinterest{The authors declare no conflict of interest.}
 
-%We may think we can use our intuition to guide us but it is more likely that it will lead us astray.
-Our ideas of conserved charges, symmetries, quantum numbers and the like may not apply in those out-of-equilibrium states.
-As we know that, during those processes, particles can transform from one type to another, it would make sense to assume they are all made of the same substance and that particles are more or less stable configurations of that substance.
-If that is the case, mass, charge and other quantum numbers may simply be the properties of those stable configurations that have no meaning in between, much like pressure or temperature for an out-of-equilibrium volume of gas.
-Even the symmetries would be qualifying the homogeneity of the equilibrium states.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
 
-%Unfortunately, it is probably worse than that.
-Suppose we want a theory that works at the smallest time resolution, at Planck scale.
-Then our set of processes $\mathcal{P}$ would correspond to those processes bounded by that time scale.
-Those will correspond to the basic building blocks with no further refinement possible.
-If we now want the law of evolution $\mathcal{E}$ as well, the time scale of $\mathcal{E}$ will be the time scale of the processes $\mathcal{P}$: we cannot see time evolution $\mathcal{E}$ as a continuous sequence of processes $\mathcal{p}_t$.
-As we can no longer pretend that between two fast time scale processes there is always another one, we now must face the fact that these may be running in different regions that may or may not overlap in space and time.
-That is, the execution of there processes is not strictly sequential and therefore not linearly ordered, in which case time cannot be modeled by a number.
+%=====================================
+% References, variant A: internal bibliography
+%=====================================
+\reftitle{References}
+\begin{thebibliography}{999}
 
-%If Hilbert spaces are not suitable to characterize states, or if real and rational numbers are not suitable to characterize time, it would mean that we may be very very far from even having the correct mathematical tools to be able to characterize those regimes.
+\bibitem[Carcassi(2018)]{Carc1}
+G.
+Carcassi, C.
+A.
+Aidala, D.
+J.
+Baker, L.
+Bieri; From physical assumptions to classical and quantum {Hamiltonian} and {Lagrangian} particle mechanics.
+{\em J.
+Phys.
+Comm.} {\bf 2018}, {\em 2(4)}, 045026.
 
-\bibliography{bibliography}
+\bibitem[Genovese(2010)]{Genovese}
+M. Genovese; Interpretations of quantum mechanics and measurement problem. {\em Adv. Sci. Lett.} {\bf 2010}, {\em 3}, 249--258.
 
+\bibitem[Bassi(2013)]{Bassi}
+A. Bassi, K. Lochan, S. Satin, T. P. Singh, and H. Ulbricht; Models of wave-function collapse, underlying theories, and experimental tests. {\em Rev. Mod. Phys.} {\bf 2013}, {\em 85}, 471.
+
+\bibitem[Daneri(1966)]{Daneri}
+A. Daneri, A. Loinger and G. M. Prosperi; Further remarks on the relations between statistical mechanics and quantum theory of measurement. {\em Nuovo Cimento B} {\bf 1966}, {\em 44}, 119--128.
+
+\bibitem[Merlin(2015)]{Merlin}
+R. Merlin; A Heuristic Approach to the Quantum Measurement Problem: How to Distinguish Particle Detectors from Ordinary Objects. {\em Int. J. Mod. Phys. B} {\bf 2015}, {\em 29(22)}.
+
+\bibitem[Tamir(2013)]{Tamir}
+B, Tamir, E. Cohen; Introduction to Weak Measurements and Weak Values. {\em Quanta} {\bf 2013}, {\em 2}, 7–17.
+
+\bibitem[Carcassi(2019)]{Carc3}
+G. Carcassi, C. A. Aidala; Assumptions of Physics. {\em In preparation} {\bf 2019}, http://assumptionsofphysics/book.
+
+\end{thebibliography}
+
+% The following MDPI journals use author-date citation: Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.
+For those journals, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%=====================================
+% References, variant B: external bibliography
+%=====================================
+%\externalbibliography{yes}
+%\bibliography{your_external_BibTeX_file}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\sampleavailability{Samples of the compounds ...... are available from the authors.}
+
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \end{document}

### 2019-05-22 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Revision submitted to BJPS
Submitted to Studies
Adding section on physicality of information entropy

@@
-\documentclass[11pt]{article}
+\documentclass[11pt]{elsarticle}
 
@@
 \title{Hamiltonian mechanics is conservation of information entropy}
-%\author{Gabriele Carcassi, Christine A. Aidala}
-
-%\ifjournal
-%	% As the page margins are so large, the full title does not fit
-%	\titlerunning{From physical principles to classical and quantum particle mechanics}
-%\fi
-
-%\date{\today}
-
-\maketitle
 
@@
 This should be at least of interest to that part of the philosophy community interested in the foundations of classical mechanics.
-Second because information entropy has a deep connection to the foundations of thermodynamics, therefore the strong link we establish here between information entropy and Hamiltonian mechanics should be of interest to an even wider audience.
+Secondly, information entropy seems to have a connection to the foundations of thermodynamics and statistical mechanics, in the sense that it is used successfully in physics practice, even on systems far from equilibrium (cite the 2 papers that Horowitz gave us).
+Therefore the strong link we establish here between information entropy and Hamiltonian mechanics should be of interest to an even wider audience.
 
@@
 
-%Now, the fact that we can use a tangent bundle to track our particle states does \emph{not} mean that conjugate momentum is a one form, in the sense of a map from a vector to a number.
-The state space is just a manifold, a set of points identified by $n$ continuous variable, and the structure it has comes from how units transform, not because the variables identify separate objects.
-Unfortunately mathematical structures strip out the units, the knowledge of the double role of coordinates is lost and then a different mathematical concept is used, one that has the same mathematical features of our particle state space but an altogether different meaning and construction.
-Changing state variables from position and momentum to position and velocity is not, as one may mistakenly understand from the math, a change of objects from a cotangent bundle to a tangent bundle: it is relabeling the same states much like changing spatial coordinates relabels positions.
-This change of variables, though, is not in general a canonical transformation which means the density will not correspond to the invariant one.
-But it is still a valid change of state variables and in fact, giving that velocity is gauge invariant, a very useful one.
-It allows to write Poisson brackets and find invariants of motion that are readily physically meaningful~\cite{Holten}.
-
-%We can summ
-
-% In fact, if $(q^1, p_1)$ and $(q^2, p_2)$ are two degrees of freedom and $\rho_1(q^1, p_1)$ and $\rho_2(q^2, p_2)$ are the respective marginal distributions, we have now four ways of characterizing that they are independent.
-%These characterizations are equivalent because they repre
-
-%This means that relationships between the entropy of the total and marginal distributions are preserved.
-For example, saying that the marginal distributions $\rho_1(q^1, p_1)$ and $\rho_2(q^2, p_2)$ are independent is equivalent to saying that the entropy of the joint distribution is equal to the sum of the entropy of the marginals: $I(\rho(q^1,q^2, p_1, p_2)) = I(\rho_1(q^1, p_1)) +I(\rho_2(q^2, p_2))$.\footnote{This is also equivalent to saying that the joint distribution is the product of the marginals $\rho(q^1,q^2, p_1, p_2) = \rho_1(q^1, p_1)\rho_2(q^2, p_2)$.} That is, giving both values is equivalent to giving one and then the other.
-This means we can think of these relationships and quantities as real physical properties assigned to the degree of freedom and not just to the coordinates.
-
-%\begin{description}
-%	\item \emph{Coordinate independence}. The unit of $q^1$ can be chosen independently of $q^2$.
-%	\item \emph{State count}. If $dq^1 dp_1$ and $dq^2 dp_2$ represent how many possible states the count of states in each degree of freedom then the total count across is the product 
-%	\item \emph{Statistical independence}. The joint distribution is the product of the marginals: $\rho(q^1,q^2, p_1, p_2) = \rho_1(q^1, p_1)\rho_2(q^2, p_2)$.
-%	\item \emph{Information independence}.
-The information entropy of the joint distribution is the sum of the information entropy of the marginals: $I(\rho(q^1,q^2, p_1, p_2)) = I(\rho_1(q^1, p_1)) +I(\rho_2(q^2, p_2))$.
-Intuitively this means that giving the value of all state variables is equivalent to giving the values of both degree of freedom one after the other.
-%\end{description}
-%These characterizations are equivalent: you can't have one without the others.
-
 \subsection*{The geometry of phase space}
@@
 
+\subsection*{On the physicality of information entropy}
+
+We want to stress, again, that in general information entropy cannot be taken as an objective physical quantity.
+It depends on the choice of symbols one uses for encoding: using letters of the alphabet or words in the English language will yield a different information entropy for the same book.
+It will depend on the choice of ensemble: in game theory one uses credence distributions that are subjective objects and therefore the information entropy will be different for different actors.
+The nature of the information entropy, then, depends on the nature of the elements, the ensemble and the values over which the distribution is defined.
+
+Moreover, we should note that our construction is not the one typically used in classical statistical mechanics.
+In that case, the elements are points across $N$ copies of phase space (one set of position and momentum for each particle) and the distribution over them is a probability distribution.
+In this case one has to be careful and show how the distribution is uniquely defined by objective physical entities.
+If one interprets said distribution as credence, like (cite Janes), then one has obvious problems.
+
+Our case, then, is by construction very particular.
+The ensemble is the state of a physical system.
+The elements are the physical subdivisions of said system, the particles.
+The set of values over which the distribution is defined is the state space of the parts.
+The densities relate what fraction of the whole can be found in a particular particle state.
+Densities and areas in terms of $q^i$ and $k_i$ are invariant pure numbers, independent of the choice of unit.
+No ``partitioning" or ``averaging" of phase space is invoked.
+The information entropy depends on no other external factors: it is an invariant state variable of the ensemble.
+It is a physical property of the system.
+
+Our statistical distribution does not correspond to the probability distribution of statistical mechanics.
+Conceptually, it corresponds to a point across the $N$ copies of phase space, a microstate, in the following way: pick a very narrow unitary distribution $\delta(q,p)$; make a copy for each particle so that it is centered around its position and momentum; sum them all.\footnote{Note that if we permute the particles we obtain the same distribution, so this repackaging does not suffer from overcounting.} That is:
+\begin{equation}
+\rho(q,p) = \sum\limits_{a=1}^N \delta(q - q_a, p - p_a)
+\end{equation}
+To form a complete link to statistical mechanics we would have to define what the probability distribution is.
+To form a complete connection to thermodynamics we would have to introduce a link to thermodynamic entropy.
+This will be outside of the scope of this work.
+At this point we simply want to show how information entropy applied to statistical distributions over phase space is indeed a physical quantity that is deeply linked to the notion of deterministic and reversible evolution and to Hamiltonian mechanics.
+
 \section{Discussion}
@@
 \rho = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
-I = &\ln (2\pi\sigma_q\sigma_p) + 1
+I = &\ln (2\pi e\sigma_q\sigma_p)
 \end{align*}
@@
 \begin{align*}
-\sigma_q\sigma_p \geq \exp (I - 1) / 2 \pi 
+\sigma_q\sigma_p \geq \exp (I) / 2 \pi e 
 \end{align*}
@@
 \begin{align*}
-\frac{\hbar}{2} &= \frac{h}{4\pi} \equiv \frac{e^{I_0 - 1}}{2 \pi} =  \frac{e^{I_0}}{2 e \pi} \leq \sigma_q\sigma_p
+\frac{\hbar}{2} &= \frac{h}{4\pi} \equiv  \frac{e^{I_0}}{2 e \pi} \leq \sigma_q\sigma_p
 \end{align*}
@@
 
-%These results represent one set of insights from our larger project, Assumptions of Physics, that aims to rederive the standard laws from a handful of physical starting points.
-
 This approach may point to a reorganization of fundamental physics on a more physically rigorous conceptual footing and more meaningful mathematical grounds which we believe is not only possible, but long overdue.

### 2019-06-11 Articles/2019-ForQuantumOptics/MDPI_template/template.tex
Submitted version
Resubmission

@@
 
+\newcommand{\edit}[1] {\emph{*** #1 ***}}
+
 %% MDPI internal command: uncomment if new journal that already uses continuous page numbers 
@@
 In this light, the different theories are specializations of a broader model.
-As unitary evolution corresponds to quasistatic evolution, violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
+As unitary evolution corresponds to quasistatic evolution, violations of the Schr\"{o}dinger equation, then, should be expected only for processes that cannot be considered quasistatic.
 }
@@
 gives the same output no matter how many times it is applied) and is therefore mathematically modeled by a projection operation on the space of statistical distributions.
-We identify quantum projections as these processes.
-If we want a deterministic law, then, it must be a quasistatic evolution in terms of those faster scale processes.
+We identify quantum projections as these processes.\edit{Quantum projections, then, should be considered in the same spirit as out-of-equilibrium thermodynamic processes: the dynamics is taken as a black box with no further characterization, there is only a clear initial and final state, and the process can, in general, increase entropy.
+Quantum states and projections are taken as prime notions in the same way that equilibrium is a prime notion in thermodynamics.} If we want a deterministic law, then, it must be a quasistatic evolution in terms of those faster scale processes.
 As the change must be small, this leads to continuous evolution over the state space and unitary evolution in the space of statistical distributions.
@@
 This will illustrate how looking at the foundations of all physical theories, and not at quantum mechanics in isolation, fosters a better general understanding of fundamental concepts.
+
+\edit{It should be noted that while the objective is to put together a general mathematical model that can fit not only quantum mechanics, but other physical theories as well, the model itself cannot be proven mathematically, in the same way that the axioms of set theory, or any other formal system, cannot be proven.
+Moreover the mathematical structure we want to impose is necessarily ``simple" because simpler structures are the more general ones (i.e.
+every mathematical object is an element of a set, the simplest structure).
+Therefore the bulk of the work is, appropriately, to present a physical justification as to why the simple mathematical structure we lay out is necessary.}
 
@@
 We will then recover unitary evolution (i.e.
-the Schroedinger equation) as the quasistatic deterministic evolution given by continuous projections.
+the Schr\"{o}dinger equation) as the quasistatic deterministic evolution given by continuous projections.
 To do so, we will introduce a different mathematical notation which will make explicit the difference between pure states, statistical distributions and mixed states, to help dispel potential confusion.
@@
 \end{equation}
-This requires the evolution to be unitary and we can use standard techniques from here, which will lead to the Schroedinger equation.
-In other words, \textbf{the Schroedinger equation is the quasistatic deterministic evolution of a process of continuous projections}.
-This approach, therefore, proceeds in the opposite way of others: instead of assuming the Schroedinger equation is always valid and deriving projections as a special case, we assume projections are valid approximations of faster scale processes and the Schroedinger equation is the special quasistatic deterministic case.
+This requires the evolution to be unitary and we can use standard techniques from here, which will lead to the Schr\"{o}dinger equation.
+In other words, \textbf{the Schr\"{o}dinger equation is the quasistatic deterministic evolution of a process of continuous projections}.
+This approach, therefore, proceeds in the opposite way of others: instead of assuming the Schr\"{o}dinger equation is always valid and deriving projections as a special case, we assume projections are valid approximations of faster scale processes and the Schr\"{o}dinger equation is the special quasistatic deterministic case.
 
@@
 As such, this approach does not technically pick an interpretation: it merely states that whatever happens during a measurement in a given interpretation is also happening at every instant in time.}
-
-%since the fast out-of-equilibrium process cannot be described with a quasi-static evolution.
-
-%The connection to measurement is simply that, as we will see later in more detail, such processes are always required if we want to measure a quantity, even in classical mechanics or thermodynamics.
-For example, if we want to measure the temperature of a gas it needs to be in equilibrium for long enough for us to conduct the measurement.
-If we are measuring the position of a ball, we need to integrate over a (small) period of time to integrate the giggling created by the random collisions of air particles.
-That is: if we want to measure the vertical spin component of an electron, something needs to make sure that we actually have and electron and its vertical component is well defined.
-So we need a process whose outputs are equilibria with those properties.
 
@@
 We could study the atoms of the molecules, yet we would have to assume that the atoms are stable under nuclear fission.
-We could study the electrons and nucleons of the atoms, yet we would have to assume they are stable under weak nuclear forces.
-As we move to a finer and finer description, we have faster and faster scale processes that we assume do not change the basic constituents.
+We could study the electrons and nucleons of the atoms, yet we would have to assume they are stable under weak nuclear forces.\footnote{\edit{This hierarchy of time scales is indeed used in many areas of science and engineering under the practice of multiscale modeling.
+\cite{Horstemeyer,Weinan} It allows one to break down a complex problem so that the relevant dynamics is studied efficiently at the right level of scale, gaining more insight and leading to faster calculations.
+Each scale will use different models and techniques, such as continuum mechanics, Monte-Carlo methods for statistical physics, molecular mechanics, quantum chemistry and so on.
+\cite{Brandt} During a computation cycle, each level of scale treats the faster and slower dynamics as fixed parameters, which are updated appropriately for the following iteration.}} As we move to a finer and finer description, we have faster and faster scale processes that we assume do not change the basic constituents.
 At each level of the description what is assumed to be in equilibrium changes and what the laws describe changes as well.
@@
 States are always equilibria of faster scale processes, which in quantum mechanics are modeled with the projection.
-States are chosen to satisfy a deterministic law of evolution, necessarily describing a quasistatic process, which in quantum mechanics is modeled by the Schroedinger equation.
+States are chosen to satisfy a deterministic law of evolution, necessarily describing a quasistatic process, which in quantum mechanics is modeled by the Schr\"{o}dinger equation.
 This general framework only uses standard physics concepts and puts no special role on the observer except the choice of what processes the system will be subjected to.
 
-Violations of the Schroedinger equation should be expected only for processes that cannot be considered quasistatic.
-If we were to implement a measurement process, then, in a way that the transfer of information between the system and the probe happens in a slow and controlled fashion, this could not provide a violation of the Schroedinger equation, nor would it tell us that all projections are quasistatic processes.
+Violations of the Schr\"{o}dinger equation should be expected only for processes that cannot be considered quasistatic.
+If we were to implement a measurement process, then, in a way that the transfer of information between the system and the probe happens in a slow and controlled fashion, this could not provide a violation of the Schr\"{o}dinger equation, nor would it tell us that all projections are quasistatic processes.
 On the other hand, a better description of the faster scale processes given by the projections cannot be given in terms of unitary evolution, as this represents the quasistatic case.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
-The following statements should be used ``conceptualization, X.X.
-and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
-and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing--original draft preparation, X.X.; writing--review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
-Authorship must be limited to those who have contributed substantially to the work reported.}
+\authorcontributions{Writing – original draft, Gabriele Carcassi and Christine A. Aidala.}
 
@@
 A. Bassi, K. Lochan, S. Satin, T. P. Singh, and H. Ulbricht; Models of wave-function collapse, underlying theories, and experimental tests. {\em Rev. Mod. Phys.} {\bf 2013}, {\em 85}, 471.
+
+\bibitem[Weinan(2011)]{Weinan}
+E. Weinan; Principles of Multiscale Modeling. {\em Princeton University} {\bf 2011}.
+
+\bibitem[Horstemeyer(2009)]{Horstemeyer}
+M.
+F.
+Horstemeyer; Multiscale Modeling: A Review.
+In Leszczyński, Jerzy; Shukla, Manoj K.
+(eds.).
+Practical Aspects of Computational Chemistry: Methods, Concepts and Applications.
+pp.
+87–135.
+ISBN 978-90-481-2687-3.
+
+\bibitem[Brandt(2002)]{Brandt}
+Brandt A.
+(2002) Multiscale Scientific Computation: Review 2001.
+In: Barth T.J., Chan T., Haimes R.
+(eds) Multiscale and Multiresolution Methods.
+Lecture Notes in Computational Science and Engineering, vol 20.
+Springer, Berlin, Heidelberg
 

### 2019-06-12 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Reviewed part about physicality of information

@@
 This should be at least of interest to that part of the philosophy community interested in the foundations of classical mechanics.
-Secondly, information entropy seems to have a connection to the foundations of thermodynamics and statistical mechanics, in the sense that it is used successfully in physics practice, even on systems far from equilibrium (cite the 2 papers that Horowitz gave us).
+Secondly, information entropy seems to have a connection to the foundations of thermodynamics and statistical mechanics, in the sense that it is used successfully in physics practice, even on systems far from equilibrium (see \cite{Jarzynski, Maes, Horowitz}).
 Therefore the strong link we establish here between information entropy and Hamiltonian mechanics should be of interest to an even wider audience.
@@
 In this case one has to be careful and show how the distribution is uniquely defined by objective physical entities.
-If one interprets said distribution as credence, like (cite Janes), then one has obvious problems.
+If one interprets said distribution as credence, like \cite{Jaynes}, then one has obvious problems.
 
@@
 The elements are the physical subdivisions of said system, the particles.
-The set of values over which the distribution is defined is the state space of the parts.
+The set of values over which the distribution is defined is the state space of the particles.
 The densities relate what fraction of the whole can be found in a particular particle state.
@@
 To form a complete connection to thermodynamics we would have to introduce a link to thermodynamic entropy.
-This will be outside of the scope of this work.
-At this point we simply want to show how information entropy applied to statistical distributions over phase space is indeed a physical quantity that is deeply linked to the notion of deterministic and reversible evolution and to Hamiltonian mechanics.
+These additional steps will be outside of the scope of this work.
+In the present work we simply want to show how information entropy applied to physical distributions over phase space is indeed a physical quantity that is deeply linked to the notion of deterministic and reversible evolution and to Hamiltonian mechanics.
 

### 2019-06-12 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Add discussion on point particles

@@
 
+\subsection*{Point particles vs infinitesimal parts}
+
+At this point one may still argue that point particles and infinitesimal parts are both abstract concepts, both factually incorrect (i.e.
+planets are neither points nor made of infinitesimal parts), so why should we prefer one concept over the other?
+
+Proper understanding of a physical theory means answering one question: why does the theory work?
+And the answer cannot be: because it does experimentally.
+The problem is not that the answer begs the question; it is that no theory works experimentally for all possible systems in all possible settings.
+In this sense, there are no ``right" theories, only ones that agree with experiments in a set of circumstances.
+The question, then, should be really taken to mean: what are the conditions under which the theory works?
+What are the assumptions or abstractions implicit in the model and where do they apply?
+
+To us, this is not an abstract intellectual question.
+It is a practical matter: I have a system to study and have a choice of different models (e.g.
+Hamiltonian mechanics, Newtonian mechanics, quantum mechanics, Markov processes, ...), how do I know which one to pick?
+
+In this light, if Hamiltonian mechanics is characterized as describing point-like objects, whose state is described by position and momentum, and that move according to a specified set of equations, it is not clear why it would apply to a planet and to a speck of dust in a vacuum, but not to a speck of dust in a fluid or to an electron.
+The size does not seem to matter and it is not clear why those equations are important: why should a point-particle be constrained to follow those equations?
+
+If we instead characterize Hamiltonian mechanics as the deterministic and reversible motion of objects reducible to infinitesimal parts, we understand when the model choice is appropriate.
+Can we pretend that the system we are studying, under the conditions we are studying it, is made of infinitesimal parts that can be studied independently?
+In the case of an electron the answer is no.
+In the case of a planet or a speck of dust, if we are studying only the overall motion, then yes.
+Can we pretend that the evolution is deterministic and reversible?
+In the case of the speck of dust in a fluid, no.
+For a planet and speck of dust in a vacuum, yes.
+
+The characterization of classical mechanics as describing point-particles, then, fails spectacularly at answering the most basic question: what is a classical system?
+Because knowing what a classical system is means knowing when classical mechanics is applicable.
+And saying that classical mechanics describes point-particles does not clarify when classical mechanics is applicable.
+Our characterization, instead, is very straightforward: a classical system is one that can be considered to be infinitesimally reducible.
+It is furthermore Hamiltonian if the evolution is deterministic and reversible.\footnote{In the same vein, \cite{AoPPhy1} finds that it is Newtonian if it is purely kinematic (i.e.
+spatial motion identify the state), it is Lagrangian if it is both Hamiltonian and Newtonian and so on.}
+
+Also note that the approach presented leads to both the structure of phase space and the equations of motion.
+There is no freedom to choose other structures.
+With point-particles, instead, it is not clear why, for example, position is not sufficient to specify the state, or why trajectories that do not obey Hamilton's equations must be discarded.
+Therefore the point-particle characterization has far less explanatory power.
+
 \subsection*{A dictionary}

### 2019-06-17 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Revised introduction

@@
 The first is that they concentrate on the mathematical structure which, unfortunately, is not enough to characterize a physical theory.
-They all note how the same physical system can be described by different mathematical frameworks.
-For example, a classical system can be characterized by position and velocity in the Lagrangian framework and by position and momentum in the Hamiltonian framework.
-They do not note, though, that the converse is also true: the same mathematical equation can represent two different physical systems.
+By mathematical structure here we mean the part of the theory that is specified by formal relationships among abstract symbols, while by physical content we mean the link, necessarily not formal, between those symbols and actual physical concepts which, ultimately, must connect with experimental verification.
+The problem is that the same physical content can be described by different mathematical frameworks.
+For example, as it is noted in all those works, a classical system can be characterized by position and velocity in the Lagrangian framework and by position and momentum in the Hamiltonian framework.
+Conversely, the same mathematical structure can be attached to different physical content.
 For example, the equation $ma + bv = 0$ in the context of Newtonian mechanics can either represent a body in an inertial frame under linear drag or a body in a non-inertial frame with no forces acting on it.
 If $t_{in}$ is time in the inertial frame, the time $t$ in the non-inertial frame is such that  $\frac{dt}{dt_{in}} = e^{\frac{b}{m}t}$.
-Mathematical equivalence does not imply physical equivalence, which is especially problematic for~\cite{Barrett2} since it focuses on the former; for similar arguments see~\cite{Coffey}.
 In other words, we can have the same physics with different math and the same math with different physics.
 The connection between math and physics is many-to-many because mathematics captures only the syntax of a particular description without its semantics, which at that point cannot be recovered.
+Therefore, if we want to analyze, compare or just understand a physical theory, it is not enough to understand the mathematical structure: we need a clear dictionary as to what each mathematical symbol ultimately represents in the physical model.
+That is, we should be able to ask simple questions such as: what does the numeric value of the Lagrangian for a given position and velocity represent?
+What does the numeric value of the symplectic form for two vectors represent?
 
-As we do not have a one-to-one map between math and physics, fixating on the mathematical structure may even confuse us.
+If we use category theory techniques, as in \cite{Barrett2}, to compare two physical theories, we cannot do anything else but compare the formal aspects.
+We may find that the formal parts of theories are inequivalent, but it could be that, once we add all the physical content that is needed to connect the formal part to actual physical systems and their measurement, they turn out to be equivalent; for similar arguments see~\cite{Coffey}.
+Or we may compare the formal structures of symplectic manifolds and Riemannian manifolds, and conclude, as in \cite{North}, that Hamiltonian mechanics is more fundamental than Lagrangian mechanics because defining only a volume element constrains the space less than defining only a line element.
+But, again, without having discussed exactly how that structure maps to the actual physical objects we cannot be sure of that conclusion.
+For example, kinetic momentum $mv$ is what is actually measured since conjugate momentum $p$ is not a gauge-invariant quantity.
+So it may be that, once we reconnect the formal theory to the actual physically defined quantities, we are forced to implicitly reintroduce the structure we appeared to be missing.
+
+Fixating on the mathematical structure without fully developing the link to the physical model it represents can also lead to confusion.
 For example, it is known~\cite{AllHamFreeParticle} that any Hamiltonian system is, at least locally, equivalent to a free particle.
@@
 Does that mean that the Hamiltonian itself is not fundamental because it is not invariant under canonical transformations?
-Does it mean that all Hamiltonian systems are the same system, since there is always a set of canonical coordinates in which $H=p^2/2m$?
+Does it mean that all Hamiltonian systems are the same system, since locally there is always a set of canonical coordinates in which $H=p^2/2m$?
 It is also known~\cite{AllSystemsAreHam} that for any first-order system of ordinary differential equations we can find an appropriate symplectic structure such that the system can be written as a Hamiltonian system.
@@
 
-This leads us to the second problem we believe the previous works may have: when trying to attach meaning the thinking and intuition is still Newtonian.
+This leads us to the second problem we believe the previous works may have: when trying to attach meaning, the thinking and intuition is still Newtonian.
 This is particularly evident in~\cite{Curiel} where physical ideas provide the starting point to get to Lagrangian dynamics.
@@
 Unfortunately this is only true in an inertial frame.
-As~\cite{North} notes, both Hamiltonian and Lagrangian mechanics are coordinate independent: the equations are valid in all reference frames even non-inertial ones.
+As~\cite{North} notes, both Hamiltonian and Lagrangian mechanics are coordinate independent: the equations are valid in all reference frames, even non-inertial ones.
 In fact, that is probably the most notable advantage of the frameworks, that fictitious forces are automatically handled.
 But none of the Newtonian concepts are coordinate independent: inertial frames, absence of forces and even conservation of energy are all frame dependent.
-This, again, leads to confusion and we may miss a key element: in the transition from Newtonian to Lagrangian/Hamiltonian mechanics we expand the reference frames in which the equations are valid (from inertial frames to all frames) at the expense of restricting the class of systems we can describe (e.g.
+This, again, leads to confusion and we may miss a key element: in the transition from Newtonian to Lagrangian/Hamiltonian mechanics, ~\cite{Curiel} must necessarily expand the reference frames in which the equations are valid (from inertial frames to all frames) at the expense of restricting the class of systems we can describe (e.g.
 dissipative forces in general require a modification of the Euler-Lagrange equation typically using the Rayleigh dissipation function).
+It is not clear when and how that transition between models happens.
 The point is, Newtonian concepts may prevent us from fully understanding Hamiltonian and Lagrangian mechanics.
 
-It should be clear that we do not fault the authors.
-Even just a century ago, there was a push in physics to find fundamental principles or laws that could serve as a starting point for the different theories.
+This attitude of treating mathematical structures like they are, by themselves, fundamental objects in a physical theory, instead of tools to capture its formal aspects, is now somewhat pervasive in some areas of physics, but it should be noted that it was not always like that.
+Even just a century ago, there was a push to find fundamental principles or laws expressed in physical terms that could serve as a starting point for the different theories.
 Newtonian mechanics, thermodynamics and special relativity serve as good examples.
@@
 
-It is only natural that, given the circumstances, a philosopher would use the only thing that seems to be well defined, the math, and try to draw understanding from there.
-But here lies the fundamental problem: that math is not created by physicists to do physics, it is created by mathematicians to do math.
+It is only natural that, given the circumstances, one would use the only thing that seems to be well defined, the math, and try to draw understanding from there.
+But here lies the fundamental problem: modern mathematics is not created by physicists to do physics, it is created by mathematicians to do math.
 The definitions of topology, differentiable manifold, symplectic manifold, Hilbert space and so on are chosen either because they are convenient to prove theorems or because they are convenient to perform calculations.\footnote{In fact, it is common in math to have two such definitions and then prove they are equivalent.} We should not expect them to be a good match for definitions that would be physically meaningful.
 In other words, there is a very good chance that looking at just the math may lead us, like horses with blinders, in a well-defined direction which is not the physically meaningful one.
-To be clear, we do not fault the mathematicians either as they are not responsible for whether their structures are used appropriately or not within physics.
+To be clear, we do not fault the mathematicians as they are not responsible for whether their structures are used appropriately or not within physics.
 

### 2019-06-21 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Adding comments from Josh

@@
 The first is that they concentrate on the mathematical structure which, unfortunately, is not enough to characterize a physical theory.
-By mathematical structure here we mean the part of the theory that is specified by formal relationships among abstract symbols, while by physical content we mean the link, necessarily not formal, between those symbols and actual physical concepts which, ultimately, must connect with experimental verification.
+By mathematical structure here we mean the part of the theory that is specified by formal relationships among abstract symbols, while by physical content we mean the link, necessarily not formal, between those symbols and actual physical concepts which, ultimately, must connect with experimental verification at least potentially.
 The problem is that the same physical content can be described by different mathematical frameworks.
-For example, as it is noted in all those works, a classical system can be characterized by position and velocity in the Lagrangian framework and by position and momentum in the Hamiltonian framework.
+For example, as it is noted in~\cite{North,Curiel,Barrett1,Barrett2}, the state of many classical systems can be characterized by a point in a tangent bundle (through position and velocity) in the Lagrangian framework and by a point in a cotangent bundle (through position and momentum) in the Hamiltonian framework.
 Conversely, the same mathematical structure can be attached to different physical content.
@@
 
-If we use category theory techniques, as in \cite{Barrett2}, to compare two physical theories, we cannot do anything else but compare the formal aspects.
+If we just use category theory techniques, as in \cite{Barrett2}, to compare two physical theories, we cannot do anything else but compare the formal aspects.
 We may find that the formal parts of theories are inequivalent, but it could be that, once we add all the physical content that is needed to connect the formal part to actual physical systems and their measurement, they turn out to be equivalent; for similar arguments see~\cite{Coffey}.
@@
 Fixating on the mathematical structure without fully developing the link to the physical model it represents can also lead to confusion.
-For example, it is known~\cite{AllHamFreeParticle} that any Hamiltonian system is, at least locally, equivalent to a free particle.
+For example, it is known~\cite{AllHamFreeParticle} that any Hamiltonian system is, at least locally, mathematically equivalent to a free particle.
 That is, we can always find a canonical transformation (i.e.
@@
 
-This attitude of treating mathematical structures like they are, by themselves, fundamental objects in a physical theory, instead of tools to capture its formal aspects, is now somewhat pervasive in some areas of physics, but it should be noted that it was not always like that.
+This attitude of treating mathematical structures in isolation as fundamental objects in a physical theory, instead of tools to capture its formal aspects, is now somewhat pervasive in some areas of physics, but it should be noted that it was not always like that.
 Even just a century ago, there was a push to find fundamental principles or laws expressed in physical terms that could serve as a starting point for the different theories.
 Newtonian mechanics, thermodynamics and special relativity serve as good examples.
-Nowadays, unfortunately, many of the current fundamental theories, including Lagrangian, Hamiltonian, quantum mechanics and quantum field theory take as their starting point a particular mathematical structure.
+Nowadays, unfortunately, many of the current fundamental theories, including Lagrangian mechanics, Hamiltonian mechanics, quantum mechanics and quantum field theory take as their starting point a particular mathematical structure.
 That is, we do not know what the value of a Lagrangian means or what the symplectic form represents; we just take them as given.
@@
 It depends on the choice of symbols one uses for encoding: using letters of the alphabet or words in the English language will yield a different information entropy for the same book.
-It will depend on the choice of ensemble: in game theory one uses credence distributions that are subjective objects and therefore the information entropy will be different for different actors.
+It will depend on the choice of ensemble: in game theory one uses credence distributions that depend on the agent and therefore the information entropy will be different for different actors.
 The nature of the information entropy, then, depends on the nature of the elements, the ensemble and the values over which the distribution is defined.
@@
 In this case one has to be careful and show how the distribution is uniquely defined by objective physical entities.
-If one interprets said distribution as credence, like \cite{Jaynes}, then one has obvious problems.
+If one interprets said distribution as credence, like \cite{Jaynes}, then one has obvious problems with viewing information entropy as a physical quantity.
 
@@
 The elements are the physical subdivisions of said system, the particles.
+The subdivisions are dictated by what parts can be studied independently.
 The set of values over which the distribution is defined is the state space of the particles.
@@
 To form a complete connection to thermodynamics we would have to introduce a link to thermodynamic entropy.
-These additional steps will be outside of the scope of this work.
-In the present work we simply want to show how information entropy applied to physical distributions over phase space is indeed a physical quantity that is deeply linked to the notion of deterministic and reversible evolution and to Hamiltonian mechanics.
+These additional steps are outside of the scope of this work.
+Here we simply want to show how information entropy applied to physical distributions over phase space is indeed a physical quantity that is deeply linked to the notion of deterministic and reversible evolution and to Hamiltonian mechanics.
 
@@
 
-At this point one may still argue that point particles and infinitesimal parts are both abstract concepts, both factually incorrect (i.e.
+At this point one may still argue that point particles and infinitesimal parts are both abstract concepts and both factually incorrect (i.e.
 planets are neither points nor made of infinitesimal parts), so why should we prefer one concept over the other?
@@
 The problem is not that the answer begs the question; it is that no theory works experimentally for all possible systems in all possible settings.
-In this sense, there are no ``right" theories, only ones that agree with experiments in a set of circumstances.
-The question, then, should be really taken to mean: what are the conditions under which the theory works?
+In this case, there are no ``right" theories, only ones that agree with experiments in a set of circumstances.
+The question, then, should be really taken to mean: what are all the necessary conditions under which the theory works?
 What are the assumptions or abstractions implicit in the model and where do they apply?
@@
 It is a practical matter: I have a system to study and have a choice of different models (e.g.
-Hamiltonian mechanics, Newtonian mechanics, quantum mechanics, Markov processes, ...), how do I know which one to pick?
+Hamiltonian mechanics, Newtonian mechanics, quantum mechanics, Markov processes, ...).
+How do I know which one to pick?
 
-In this light, if Hamiltonian mechanics is characterized as describing point-like objects, whose state is described by position and momentum, and that move according to a specified set of equations, it is not clear why it would apply to a planet and to a speck of dust in a vacuum, but not to a speck of dust in a fluid or to an electron.
-The size does not seem to matter and it is not clear why those equations are important: why should a point-particle be constrained to follow those equations?
+If Hamiltonian mechanics is characterized as describing point-like objects, whose state is described by position and momentum, and that move according to a specified set of equations, it is not clear why it would apply to a planet and to a speck of dust in a vacuum, but not to a speck of dust in a fluid or to an electron.
+The size does not seem to matter and it is not clear why Hamilton's equations are important: why should a point-particle be constrained to follow those equations?
 
@@
 
-Also note that the approach presented leads to both the structure of phase space and the equations of motion.
+Finally, note that the approach presented leads to both the structure of phase space and the equations of motion.
 There is no freedom to choose other structures.

### 2019-09-07 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Adding Shannon entropy first draft

\documentclass{article}

%\usepackage{arxiv}

\title{Shannon entropy as the variability of elements within a distribution}

\author{
	Julian Barbour, 
	Gabriele Carcassi
}

\begin{document}

\maketitle

\begin{abstract}
TBD
\end{abstract}

\section{Introduction}

Current disputed status of Shannon entropy as a physical object.

Immense regularity in nature leads to categorization of elements intro distributions. Stress that these distributions may be of different nature? I.e.: statistical distributions of elements that exist at the same time, probability distributions of events that may happen given similar circumstances, credence distributions of expectation for an agent that something will happen. May be better to have a another word instead of "probability" when talking about the elements of the distribution. Maybe "weight" or "fraction"?

A natural question is to ask how much variability an element has within a distribution. Tempted to just count the number of cases over which the distribution is defined. But the shape also affects the variability (e.g. a distribution concentrated in few values means less variability of the elements than one equally distributed).

In this paper we will show that $- \sum p_i \log p_i$ is the most natural form to describe the variability of the elements of a distribution.

(NOTE: defining the concept like this is actually broader than Shannon entropy, which assumes source/destination/message.)

\section{Variability within a distribution}

The general setting is the following. We have a set of elements $E = \{e_\alpha\}_{\alpha=1}^N$. These can represent the galaxies in the Laniakea Supercluster, the animals in the Gal\'{a}pagos islands, the molecules in a box of gas given a particular macrostate, the different takes of spin measurement for a given quantum state, or the words in the complete works of Shakespeare. The choice of the particular set is driven by the interest and practices of scientists and their fields of study and therefore it is, in this sense, discretional. Yet, it must be objective in at least one respect: once a choice is made the actual elements are the same for everybody. Once we arbitrarily decided we will study the galaxies in the Laniakea Supercluster, it is a matter of fact that the Milky Way will be included. This also means that, whatever properties those elements will have, they will represent a matter of fact about that set. In turn, this will make some choices arguably more appropriate or interesting than others. For example, in taxonomy the set of all animals that have feathers may be more interesting than the set of all animals that can fly: the first includes only birds while the second includes most birds, a few mammals and a lot of insects.

Once a set of elements is chosen, we select a property or a set of properties we want to use to characterize the elements. That is, we have a set of possible values $Q=\{q_i\}_{i=1}^I$ and a map $q : E \to Q$ that associates a value for each element. These can represent the galaxy types, the genus of the animal, the possible states of the molecules, the possible values for spin or the spelling of words. This will give us a sequence $\{q(e_\alpha)\}_{\alpha=1}^N$ of the descriptions that are associates with each element in the set. Again, the choice of properties is ultimately discretional, grounded on what particular aspects of the elements we are studying in a given case. Yet, like before, once a choice is made the value for each element is objectively defined. Once we arbitrarily decided we are studying galaxy types it is a matter of fact that the Milky Way is a barred spiral galaxy. Again, one may find that some choices are arguably more appropriate or interesting than others.

Having decided what elements to study and the level of description we are interested in, we bin them; that is, we group them based on that description, disregarding the identity of the particular element. What we will be interested in is only the relative frequency $p_i = N_i / N = \left| \{e_\alpha \in E | q(e_\alpha) = q_i \} \right| / N$ of the elements within each bin. We may do so because we are either not interested or not able to further distinguish the elements. Whatever the reason, once the choices of elements and properties are made, this relative frequency is objective. The nature of the $p_i$ will depend on the previous choices. It may represent the fraction of elements if the set is constituted by a group of objects. It may represent a probability of an outcome is the set is constituted by different takes of similarly prepared systems. If the set is infinite, it will represent some type of convergence. We will therefore call $p_i$ weights, to be clear that we make no commitment on whether we have fractions, frequentist probabilities, Bayesian probabilities or other notions.

Now that we have constructed our distribution, we want to construct an indicator $S$ that quantifies how much variability the elements exhibit within the distributions. That is, we want to quantify the degree of diversity that the values can have within the distribution. To that end, we want to define some suitable requirements for $S$.

We may be tempted to use standard statistical quantities, like the range or the variance, but this is not possible. First, if the values associated to the bins are non-numerical (e.g. types of galaxies, words), statistical quantities are not well defined. Second, permutation of the values does not have an impact on variability of the elements, while statistical quantities will in general be affected. This tells us that our indicator cannot depend on $Q$ itself, but only on the weights $p_i$. That is, we require that $H=H(p_i)$.\footnote{Note that for continuous quantities the weights $p_i$ are affected by the choice of $Q$. We will concentrate on this problem later.}

We  expect small changes in a distribution, small changes of the weights, to produce small changes in variability, therefore we require $H$ to be a continuous function with respect to the $p_i$. We also expect that if we have a uniform distribution $p_i = 1/I$, then greater $I$ should correspond to greater variability. Therefore, in the case of uniform distribution, we require $H$ to be monotonically increasing with $I$.

As we noted before, the level at which we describe the elements is not absolute and can change. For example, we may choose to group the animals in the Gal\'{a}pagos islands first by class (e.g. mammals, birds, reptiles) and then later refine the mammals by species. What we would like to say is that the variability of the overall distribution $H_T = H_C + p_M H_M$ will be the variability over the classes $H_C$ plus the variability of the mammals $H_M$ weighted by the fraction of mammals $p_M$.

To sum up, we have the following three requirements:
\begin{enumerate}
    \item $H$ depends only on $p_i$ and it does so continuously
    \item If $p_i=1/I$ then $H$ is a monotonically increasing function of $I$
    \item Let $p_i$ and $q_j$ be the weights for two distribution respectively over $I$ and $J$ bins. Let $r_k$ be the distribution over $K=I+J-1$ bins constructed by expanding the $a$-th bin of the first distribution using the second distribution. More specifically, let $1 \leq a \leq I$, then $r_k = \{p_1, p_2, ..., p_{a-1}, p_{a}q_1, p_{a}q_2, ..., p_{a}q_J, p_{a+1}, ..., p_I \}$. Then $H(r_k) = H(p_i) + p_{a} H(q_j)$.
\end{enumerate}

We note that these are exactly the requirements Shannon put forth for his entropy\cite{Shannon}, and that these requirements lead to only one possible option: $H(p_i) = - \sum p_i \log(p_i)$.

The Shannon entropy, then, is a clear and important attribute of a distribution: it quantifies the variability of the elements within the distribution. It quantifies the variety of values within the distribution. It is true that the choice of the elements and binning are not fixed, it is true that the meaning of the weights depends on what the distribution is describing, and therefore that the precise meaning of this variability is context dependent. But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope a probability and so on. And like a real number represents a quantity, transcending what is physically quantified, the Shannon entropy represents the variability of the elements in a distribution, transcending what the distribution physically signifies.

The Shannon entropy may represent uncertainty in some cases, if the weights are probabilities or credences, but not in the general case. If the weights portray the fraction of the elements that have a certain property, like the fraction of galaxies in the Laniakea Supercluster that are barred spirals, there is nothing uncertain about the distribution. In this case, the Shannon entropy represents how much variation we have within galaxies in terms of galaxy types.

The Shannon entropy may represent knowledge in some cases, but not in general. Consider the following two cases:
\begin{enumerate}
    \item There is 50\% chance you won one million dollars in the lottery and 50\% chance you didn't;
    \item There is 50\% chance you won one million dollars in the lottery and 50\% chance you won half a million dollars.
\end{enumerate}
The distribution in both cases is the same, two bins 50\% chance each, and so is the Shannon entropy. Yet, you know more in the second case: you know you have won at least half a million dollars.

Unfortunately, entropy in general is often associated with vague characterizations as the ones presented. It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending by the authors, which leads to confusion and misunderstanding.  The characterization of Shannon entropy we have given, on the other hand, applies to all cases, it is enough to understand the assumptions required to rederive it and therefore is more fundamental. If we look at the galaxies in our universe, what variability do they exhibit in terms of their types? If we look at the animals in the Gal\'{a}pagos island, what variability is expressed in terms of their species? If we look at the molecules in a given macrostate, what variability do they express in terms of their microstates? If we look at the words in the complete works of Shakespeare, what variability we find in its vocabulary? All these questions are interesting, all these questions are profound and all these questions are answered by a single concept: the Shannon entropy.

If the Shannon entropy is a measure of variability, why is it connected to information? What is it measuring and in what units?

\section{Units of variability}

We have seen that the Shannon entropy measures the variability within a distribution but, in the end, it is a numeric quantity and we cannot expect to understand the concept if we don't understand what the number represents and in what units.

We can gain insight with the following example. Suppose we fix a distribution, say the animals in the Galap\'{a}gos islands. Suppose I pick a specific animal from the distribution and you want to know its species. Suppose the only way for you to get that information is to ask me a series of questions. Suppose each questions has two possible answers, yes or no. How many questions would you have to ask? In other words, we are playing a game of Twenty Questions.

Not all questions will be able to extract the same amount of information. Some questions, like ``is it an animal?" would be redundant. Others, like ``is it an American Flamingo (Phoenicopterus ruber)?", would give us a lot of information in the positive case but little in the negative case. It should be clear, though, that there must be a limit to how few questions you must ask to get to the answer: a single yes/no question cannot be enough. It should, hopefully, be intuitive that to a greater variability within the distribution will correspond a greater number of questions you must ask. That is exactly what the Shannon entropy measures: the minimum average number of questions one has to ask to identify a value in the distribution. It gives us the number of questions for an ideal strategy for our game of Twenty Questions.

If we have binary questions, the logarithms will be in base two and the unit for Shannon entropy will be bits. It will indicate the average number of yes/no questions we need to identify an element within the dsitribution. In general, you can pick any base $b$: in base three we have ternary questions and trits, for ten we have questions with ten possible answers and digits. We can also pick a non-integer base, like the natural base for logarithms, and we will have nats. This is why the Shannon entropy is fundamental in information theory, because it quantifies how much information is needed to transfer a value from a distribution.

Defining a set of questions means choosing an encoding as we are choosing how the information gets codified into our series of bits. To understand how this works, we can briefly review the Hauffman coding, which is the optimal algorithm for symbol-by-symbol coding with a known probability distribution. The idea is that we want all possible answers to each question to be balanced, to provide the same amount of information. The reason is that making one answer more specific (i.e. it applies in fewer cases) means making another less specific (i.e. it applies in more cases). In the case of binary questions, then, we want the probability to answer yes or no to be 50\%.

For example, suppose we have a population of pets in a country is as follows:
\begin{itemize}
    \item dogs 27\%
    \item cats 48\%
    \item fish 10\%
    \item bird 8\%
    \item small mammal 4\%
    \item reptile 3\%
\end{itemize}
For the first question, we group cats on one side and everything else on the other, to form a 48/52 split. So we can ask ``is it a cat?". If the answer is yes, we finished. If not, we need to continue. We can group dogs on one side and everything else on the other to form a 27/25 split. So we can ask `` is it a dog?". If the answer is yes, we are done. If not, we continue. We can group fish with reptile and bird with small mammal to form a 13/14 split. So we can ask ``is it a fish or a reptile?" If the answer is yes, the followup question would be ``is it a fish?" If the answer is no, the followup question would be ``is it a bird?" With this scheme, the enconding is as follows:
\begin{itemize}
    \item dogs 27\% - 01 - 2 bits
    \item cats 48\% - 1 - 1 bit
    \item fish 10\% - 0011 - 4 bits
    \item bird 8\% - 0001 - 4 bits
    \item small mammal 4\% - 0000 - 4 bits
    \item reptile 3\% - 0010 - 4 bits
\end{itemize}
The average bit for encoding is $(.48) * 1 + (.27) * 2 + (.1 + .08 + .04 + .03) * 4 = 2.02$ bits while the Shannon entropy is 1.98 bits. That is, the coding is very close to the ideal case.

It is important, at this point, to understand that the technical use of the term information in information theory does not equate to the normal use of the term which refers to knowledge, intelligible data. The bits by themselves, the yeses and the nos, the ones and the zeros, do not provide knowledge. They need the context of the questions and the distribution to become actual information. For example, when opening a jpeg file, the file itself does not contain the instructions of how to read it. If you do not happen to know what a jpeg is and how to read it, you are not going to be able to interpret, to decode, the string of bits into actual intelligible data. The questions, the distribution, the context are considered given, communicated out-of-band through another scheme. As with any semantic content, this cannot be easily formalized and quantified.

The Shannon entropy of the distribution, then, has nothing to do with the information the distribution itself holds. The distribution is not what is being encoded. The entropy is the information needed to go from the distribution, which is given, to an individual element. It's really the information gap from the population to an element. That is why some people say the entropy is ``lack of information", because it's the information the distribution does not have. But, again, this is just missing the point of what it is being described. If one is not interested in identifying elements there is no ``missing information".

In communication and information theory, information is really encoded information. Communication systems and information processors have no idea whether the source of the data is a digital thermometer or a pressure sensor. There is no knowledge per se, just symbol manipulation that may represent different concepts in different contexts.

The term information entropy, then, is unfortunate for two reasons. First, because is not really entropy in the thermodynamic sense: it is not defined on states, it does know about irreversible processes and it is unrelated to maximization at equilibrium. In fact, it is not related to physical systems. Second, it is not really information in the general sense, only in the very narrow technical sense of coded information within a communication or information system.

So why is it so successfully used in statistical mechanics? Why does it seem to play a fundamental role even for systems far from equilibrium?

\section{Connection to statistical mechanics}

As we saw in the previous sections, Shannon entropy is a very general concept that applies to any type of distribution, no matter what the problem domain. To understand its link to the thermodynamic/statistical mechanics entropies we need two answer two questions. First, what does the Shannon entropy represents in the context of statistical mechanics? Second, why does the Shannon entropy have the same properties of the thermodynamic entropy? That is, we have to understand why it increases during irreversible evolution, it is conserved during reversible evolution, it is maximized at equilibrium and it is additive for independent systems.

The first question is straight-forward to answer. The macroscopic variables of a system (e.g. volume, pression, energy) do not give a full description of its pars. If we fix a macrostate and then measured the state of all microscopic constituents, we would not find the same answer every time. The choice of a macrostate, then, identify a set of possible microstates that are compatible with it. It is then valid to ask the question: what is the variability of microstates one can find within a given macrostates? This would be answered by the Shannon entropy, which will measure the variability in bits required to identify a microstate given the macrostate.

For the others, suppose we have a volume of gas at a particular temperature and pressure in perfect equilibrium. While the microstate is constantly changing, the macrostate remains the same. Effectively, the equilibrium is simply permutating elements of the distribution. As we saw, permutations do not change the variability which has to remain constant. At equilibrium the Shannon entropy has to remain constant.

Suppose we compress the gas adiabatically and then let it re-expand adiabatically to the original state. As the system is isolated, the whole process is deterministic and reversible. An initial microstate has to map with a specific intermidiate and final microstate. Identifying an element at each stage is equivalent to identifying an element in any state. The variability within the distribution at each step, then, cannot change: the same number of bits, the same information, must be enough to identify initial, intermediate or finale microstates. Deterministic and reversible processes conserve Shannon entropy.

Now suppose we have a process that starts from an out-of-equilibrium state and ends in an equilibrium. The initial state will not be fully described by the standard macroscopic variables. In fact, there will be many out-of-equilibrium states that will lead to the same final equilibrium. In this sense the process is irreversible because knowing the final macroscopic variables will not be enough to reconstruct the initial macroscopic state. Different possible initial out-of-equilibrium states merge into a final equilibrium state.

If we assume, though, that the microstate evolves deterministically and reversibly, knowing the final microstate would allow us to reconstruct the initial microstate and what initial out-of-equilibrium macrostate would be compatible with it. As different out-of-equilibrium macrostate merge during the evolution, they merge the possible initial conditions that lead to that same macrostate. The variability of microstates within the initial out-of-equilibrium macrostate, then, must be smaller than the variability of the microstates withing the final equilibrium state. The amount of information needed to identify a microstate at equilibrium allows us to identify the initial microstate across multiple out-of-equilibrium states and therefore is greater than the information needed to identity a microstate in an out-of-equilibrium macrostate. Irreversible processes increase the Shannon entropy which, as we saw before, stabilizes at equilibrium.

Now suppose we have two independent volume of gasses. This means that picking the microstate of one system does not constrain the microstate of the other. That is, the distribution over microstates are statistically independent. In this case, the Shannon entropy is additive.

The Shannon entropy, then, has all the right properties to establish a conceptually meaningful connection with the thermodynamic/statistical mechanics entropies. Yet, there are some mathematical details that need to be worked out. These details don't change the overall intuitive picture, but they present yet again opportunities for confusion.

The first is the role of the Boltzmann constant $k_B$, which is a conversion factor between the Shannon entropy $S = - \sum p_i \log(p_i)$ and the Gibbs entropy $S_G = - k_B p_i \sum \log(p_i)$. The issue here is that temperature, internal energy and entropy are typically linked by the relationship $T = \frac{\partial U}{\partial S}$. Historically, temperature and internal energy are the quantities that were defined first. If we measure temperature in Kelvin and energy in Joules, the above relationship forces us to measure entropy in Joules/Kelvin. If we changed units, $k_B$ would change accordingly.

However, this is not the only way to define the relationships between those quantities. Another way is to have $\beta = \frac{1}{k_B T} = \frac{\partial S}{\partial U}$. In this setting, $U$ is measured in Joules, $S$ in nats (which are dimensionless), $\beta$ is measured in inverse Joules. Temperature is still measured in Kelvin, but we do not regard it as a primary objects. We find this approach, sometimes called entropy first, prefereable for a number of reasons, expecially in statistical mechanics. First, entropy is invariant under coordinate transformations and therefore it is better to keep it a pure number and not mix it with units of energy, which is not invariant under boost.\footnote{This is essential if one ever hopes to put statistical mechanics in a relativistic setting.} Second, the maximization of entropy at equilibrium guaranties us that $S(U)$ is a single valued function (i.e. the maximum given the value of $U$) and $\frac{\partial S}{\partial U}$ is therefore well defined. On the other hand, $U(S)$ is not in general single value (e.g. in the Ising model). This means that temperature can be infinite where the curve inverts, where small changes of energy do not change the entropy, and can be negative in the other region. That is, the infinitely cold system is temperature zero, which cannot be achieved, as it warms up temperature rises until it become infinite, and as the system becomes even hotter the temperature is negative and increase toward zero, which would correspond to an infinitely hot system. On the other hand, $\beta$ is always finite and cannot be infinite. An infinitely cold system, then would correspond to positive infinite $\beta$, as the system warms up beta decreases, passes from zero, becomes negative and an infinitely hot system corresponds to minus infinite $\beta$: this is a much better property for a physically meaningful quantity. Third, when setting up entropy maximization, $\beta$ is quite naturally the Lagrange multiplier associated with energy conservation, therefore mathematically it has a much clearer role. Lastly, one can proceed in both thermodynamic and statistical mechanics with just $\beta$, and convert to temperature only when we need a value on the Kelvin scale. For these reasons, it seems to us $\beta$ plays the more fundamental role and $k_B$ is the conversion coefficient to go from $\beta$ to temperature measured in Kelvin. Adding and removing the Boltzmann constant from the entropy expression, then, is equivalent to shift from an energy/temperature first approach to an entropy/$\beta$ first approach and should be not seen as physically significant.

The other point of possible confusion is that in statistical mechanics there are actually two distributions. A single microstate is itself a distribution of particles over their particle states. There are actually $N$ particles existing at the same time, each with its own particle state. And a macrostate is a probability distribution over microstates. It is important to keep these two distributions straight, especially because a microstate can be represented in two ways.

The first way is as a distribution over a single copy of phase space, sometimes referred to as $\mu$-space. In this case the distribution represents the fraction of particles that are in that state. The Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$ is expressed in this space. On $\mu$-space, the Shannon entropy $-\int\rho \log \rho$ does not coincide with the thermodynamic entropy for a simple reason: it does not depend linearly with the number of particles. It is not an extensive quantity. The reason is straight forward: the Shannon entropy works on normalized distributions and gives us the variability of a single element. In this case, the system is composed by $N$ elements, not one. One can use the expression $-N\int\rho \log \rho$ as this represents the variability of $N$ elements.\footnote{This is actually an approximation of the multinomial distribution for large N, which would give the precise result.}

Alternatively, a microstate can be represented as a point over $N$ copies of phase space, sometimes referred to as $\Gamma$-space. It is on this space that the Gibbs entropy is defined and where thermodynamic entropy happens to coincide with the Shannon entropy. Since this is a probability distribution, the microstate corresponds to a single element within the distribution defined by the macrostate. Yet, one has to be careful of two things. The first is that permutations of particles within $\Gamma$-space lead to different points while should not lead to different microstates. This leads to the widely known problem of overcounting which needs to be addressed in the standard way. The second problem is that we should not think of the particles as literally pointlike: they would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon entropy. It is more appropriate, both mathematically and conceptually, to think of indistinguishable particles as identical distributions over $\mu$-space, each contributing the same amount of Shannon entropy.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these single particle distributions is a natural way to incorporate the effects of the uncertainty principle.}

The last point of confusion is the invariance of entropy under coordinate transformations. The use of phase-space in statistical mechanics is of paramount importance for one reason and one reason only: it is the only type of manifold where entropy can be defined in a way that is coordinate invariant. The issue is that we do not define fractions/probabilities over continuous variables: we defined densities. For a discrete case, we have a finite amount, like 3\%. For a continuous case we have an infinitesimal contribution, like 3\% per meter. If we switch labels in the discrete case, 3\% remains 3\%. If we switch to kilometers in the continuous case, we have 3000\% per kilometer. The continuous version of the Shannon entropy is affected by a coordinate transformation by the integral of the logarithm of the Jacobian determinant. Any transformation where the Jacobian determinant is not unitary, where the volumes are not preserved, does not preserve the Shannon entropy. This is not a problem for the Shannon entropy in general. It is a problem if we want to give the Shannon entropy any physical meaning.

What happens in phase space is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components. Therefore the areas $d\hat{q}^i d\hat{p}_i = dq^j dp_j$ remain the same. The Jacobian determiants are unitary and the Shannon entropy is invariant. It is not the invariance of the areas over Hamiltonian evolution that is important: it is the invariance over coordinate transformations. Invariance over time would be insignificant if we first didn't have a quantity that was observer independent.

This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\rho \log \rho$. If we use distributions over position and velocities or over momentum only, the expression we changed for those particular variables in those coordinate systems.

These technical details are important to not get confused, but they do not alter the conceptual point. The Shannon entropy and the thermodynamic entropy, though describing different things, are conceptually related. The variability of the state of the parts within the state of the whole is what the Shannon entropy quantifies if applied to the macrostates of thermodynamics and the microstates of statistical mechanics. This variability increases during irreversible evolution, more microstates converge to the same macrostate until equilibrium is reached. The variability of one system does not affect the variability of an independent one and therefore it simply sums. This is not a coincidence: this is what thermodynamic entropy ultimately is about.

\section{Conclusion}

\bibliographystyle{unsrt}  
\bibliography{references}

\end{document}

### 2019-09-13 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Adding introduction and comments (Julian)

@@
 \documentclass{article}
+\textheight 8.9 in \textwidth 6.5 in \oddsidemargin -10pt \topmargin
+-30pt
 
 %\usepackage{arxiv}
+%\usepackage[usenames,dvipsnames]{color}
 
-\title{Shannon entropy as the variability of elements within a distribution}
+\title{Universal Characterization of Variability by \\Shannon's $\sum_ip_i\,\textrm{ln}\,p_i$ Formula}
 
@@
 
+\tableofcontents
+
 \begin{abstract}
@@
 
-\section{Introduction}
+\section{Introduction\label{int}}
+
+A world without any variety is barely conceivable.
+It certainly would not admit scientific investigation.
+One of us recalls but has been unable to relocate a sentence in one of Leibniz's many writings to the effect that reality \emph{is} variety.
+This is consistent with his view that any contingent thing is defined by a collection of attributes, which necessarily must be distinguishable
+at some epistemic level.
+Paragraph 58 of Leibniz's \emph{Mondadology} asserts that the principle underlying the universe is that it should realize ``as much variety as possible, but with the greatest order possible''.
+
+Without committing ourselves to Leibniz's overarching principle, our aim in this paper is to draw attention to the specific nature of the variety that can be observed in the universe and the unique suitability of Shannon's entropy (or information) formula $\sum_ip_i\,\textrm{ln}\,p_i$ for the characterization of the variety.
+Indeed, as we will show in the next section, opportunities for application of Shannon's formula are ubiquitous both on the earth and in the universe at large.
+
+We also wish to comment on the various interpretations of Shannon's formula and attempt to answer this question: what does the formula represent?
+It is well known that when Shannon recognized the importance of his formula in communication theory he sought von Neumann's advice on the name he should give to the expression he had found.
+He was told to call it \emph{entropy}, first, because Botlzmann had used an identical formula to define entropy in statistical mechanics and, second, nobody understood entropy and that would provide protection from criticism [exact words to be found].
+We believe von Neumann's advice was not totally helpful and has given rise to confusion.
+Indeed, it is often claimed that Shannon's formula is epistemic in nature and represents ignorance.
+There certainly are situations in which that is true but equally there are others in which the use of the formula is very appropriate and no ignorance is involved at all.
+Examples will be given in the next section.
+
+This leads us to propose a different characterization of Shannon's formula that is both more general and more appropriate.
+Indeed, we will argue that the formula is a measure of the \emph{variability within a distribution of things of a given kind}.
+Moreover, the formula has a very wide range of applicability because the universe contains many different kinds of things that, through having common attributes, can be recognized as belonging together despite the fact that their attributes are expressed to different degrees or in different ways.
+To give a first example, the girls in a school can differ in the length and colour of their hair.
+More examples will follow in the next section.
+
+\section{Comments on Boltzmann's $H$ Theorem}
+
+[This is for you Gabriele].
+In Boltzmann's 1872 paper, he introduced the function that later was called the $H$ function, though it seems he meant to call it [upper case] $\eta$ and it was printed as $E$ (which somehow later became $H$).
+In the paper Boltzmann describes (in $\mu$-space) a gas of monatomic molecules that has already become spatially uniform and also has the same distribution wrt the particles' kinetic energies $x$ at each space point.
+He therefore considers a spatially constant distribution function $f(x,t)$ and wants to show that whatever its initial value it will tend in accordance with the Boltzmann equation (which he has previously derived and I won't give here) to the Maxwell distribution $f(x,t)=C\sqrt xe^{-hx}$ as $t\rightarrow\infty$.
+He defines his function $E$ as follows:
+\begin{equation}
+E=\int_0^\infty f(x,t)\left\{\textrm{log}\left [{f(x,t)\over \sqrt x}\right ]-1\right\}\textrm dx\label{hfun}
+\end{equation}
+and aims to prove that it cannot increase and attains its least value when $f(x,t)$ becomes the Maxwell distribution.
+Boltzmann does not give the argument that leads him to the function (\ref{hfun}), but somewhat later in the paper he notes that the $-1$ in the integrand merely reflects the fact that
+$$
+\int_0^\infty f(x,t)\textrm dx
+$$
+is the total number of molecules in unit volume, so that 
+\begin{equation}
+E=\int_0^\infty f(x,t)\textrm{log}\left [{f(x,t)\over \sqrt x}\right ]\textrm dx\label{hfunb}
+\end{equation}
+also cannot increase.
+
+Boltzmann does not say how he arrived at the function (\ref{hfun}), but its form is obviously dictated by the form of the Maxwell distribution.
+I'm terrible at equations, but I assume the $\sqrt x$ in the denominator of the integrand in (\ref{hfunb}) is because the usual Shannon expression is `skewed' by the subsidiary Lagrangian multiplier that fixes the total energy to a definite value.
+Later in the paper, when considering polyatomic molecules, he arrives by means of arguments I have not yet attempted to understand at the familiar $f\, \textrm{log}\,f$ form.
+
+In his very concise 1860 derivation of his energy distribution function, Maxwell commented that
+
+\begin{quote}\small
+
+It appears from this proposition that the velocities are distributed among the particles according  to the same law as the errors are distributed among the observations in the theory of the ``method of least squares''.
+
+\end{quote}\normalsize
+
+In his much more substantial 1867 paper, Maxwell noted that his 1860 derivation could be seen as `precarious' and gave an alternative proof.
+However, this does not eliminate the mathematical identity of his distribution with the error distribution.
+Would you agree that this (somewhat indirectly) establishes another example of variablity within a distribution of things [which I think we might use instead of `elements'; `members' (of a set) is another word we might consider]?
+
+\vspace{2in}
+
+TODO: Introduction and Conclusions still need to be done
 
@@
 
-\section{Variability within a distribution}
+\section{Variability within a Distribution\label{vwd}}
 
@@
 
-\section{Units of variability}
+\section{Units of Variability\label{uv}}
 
@@
 
-\section{Connection to statistical mechanics}
+\section{Connection to Statistical Mechanics\label{csm}}
 

### 2019-09-18 Articles/2019-Entropy/Source/entropy-article.tex
First notes on entropy article

\documentclass[letterpaper]{article}

\begin{document}

\title{TBD}
\author{Gabriele Carcassi, Christine A. Aidala \\ University of Michigan}

\date{\today}

\maketitle

\begin{abstract}
	TBD
\end{abstract}

\section{Introduction}

In the intro, describe only the discrete case as it simpler to understand and already contains all the physics. The continuous case adds the complication of

Evolution is a temporal sequence of states. Process is a set of possible evolutions. What we want to study is how a statement about a state at a particular time selects a particular set of possible evolutions and therefore selects a set of possible states at a previous or past time.

Define determinism and reversibility in terms of what evolutions exist. Note that a dynamical system is a deterministic system.

Note how the number of states does not always match the number evolutions it is compatible with. Note that under deterministic evolution the number of compatible evolutions can only increase. For dynamical systems, then, the number of evolutions associated to a state can only increase in time and it is stationary for equilibria.

If we compose independent systems, the number of possible states and trajectories is the product. If we take the logarithm, then those numbers are additive. Conceptually the entropy is the logarithm of the number of evolutions compatible with a given state. This article shows how this simpler and more general concept leads, in special cases, to the Boltzman entropy and the Shannon entropy. We will also see how this same concepts is applicable more in general, including the case of non-Hamiltonian systems.

\section{Different level of precision}

We regard states as points in a space. Conceptually, this presents problems. Mathematically, what we care about is the sigma algebra.

Introduce theoretical domains as point-less spaces, descriptions, that follow a countable boolean algebra. A size is given by a measure, additive map from the theoretical domain to the reals.

To define a unit system, we take one statement to be the unit, the size of a set can be defined as the ratio of the measure of the set and the measure of the set of the unit.

Two statements are independent if all minterms are possible, are not contradictions.

\bibliography{bibliography}

\end{document}

### 2019-09-18 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Adding entropy as state variable

@@
 To understand its link to the thermodynamic/statistical mechanics entropies we need two answer two questions.
-First, what does the Shannon entropy represents in the context of statistical mechanics?
+First, what does the Shannon entropy represents in the context of statistical mechanics and why is it a state variable?
 Second, why does the Shannon entropy have the same properties of the thermodynamic entropy?
@@
 If we fix a macrostate and then measured the state of all microscopic constituents, we would not find the same answer every time.
-The choice of a macrostate, then, identify a set of possible microstates that are compatible with it.
+The choice of a macrostate, then, identifies a set of possible microstates that are compatible with it.
 It is then valid to ask the question: what is the variability of microstates one can find within a given macrostates?
 This would be answered by the Shannon entropy, which will measure the variability in bits required to identify a microstate given the macrostate.
+This quantity is a property of the macrostate: it is a state variable.
 

### 2019-09-18 Articles/2019-ForQuantumOptics/article.tex
Moved all text back to article

--- 2019-04-01 Articles/2019-ForQuantumOptics/article.tex
@@
 Assuming a quasistatic evolution, then, means the equilibria of the faster scale processes will evolve slowly, giving rise to continuous and unitary evolution.
-These ideas are sufficient to get the general form of the projection postulate and the time evolution of quantum mechanics.
-Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
+In this paper we show that these ideas are sufficient to get the general form of the projection postulate and the time evolution of quantum mechanics, while also creating a model that applies to other physical theories, such as classical mechanics and thermodynamics.
+In this light, the different theories are specializations of a broader model.
+As unitary evolution corresponds to quasistatic evolution, violations of the Schr\"{o}dinger equation, then, should be expected only for processes that cannot be considered quasistatic.
 \end{abstract}
@@
 
-In an effort to better understand why the frameworks of classical and quantum mechanics are what they are, we have worked on identifying a minimal set of physical assumptions from which both can be derived.[cite] This gave us a new perspective that is somewhat unusual among people who work on fundamental questions in physics.
+In an effort to better understand why the frameworks of classical and quantum mechanics are what they are, we have worked on identifying a minimal set of physical assumptions from which both can be derived.\cite{Carc1} This gave us a new perspective that is somewhat unusual among people who work on fundamental questions in physics.
 The typical view is that the theories are so different that they require a radically different set of starting points, and that classical mechanics in particular should only be viewed as an approximation of quantum mechanics.
@@
 gives the same output no matter how many times it is applied) and is therefore mathematically modeled by a projection operation on the space of statistical distributions.
-We identify quantum projections as these processes.
+We identify quantum projections as these processes.Quantum projections, then, should be considered in the same spirit as out-of-equilibrium thermodynamic processes: the dynamics is taken as a black box with no further characterization, there is only a clear initial and final state, and the process can, in general, increase entropy.
+Quantum states and projections are taken as prime notions in the same way that equilibrium is a prime notion in thermodynamics.
 If we want a deterministic law, then, it must be a quasistatic evolution in terms of those faster scale processes.
@@
 This will illustrate how looking at the foundations of all physical theories, and not at quantum mechanics in isolation, fosters a better general understanding of fundamental concepts.
+
+It should be noted that while the objective is to put together a general mathematical model that can fit not only quantum mechanics, but other physical theories as well, the model itself cannot be proven mathematically, in the same way that the axioms of set theory, or any other formal system, cannot be proven.
+Moreover the mathematical structure we want to impose is necessarily ``simple" because simpler structures are the more general ones (i.e.
+every mathematical object is an element of a set, the simplest structure).
+Therefore the bulk of the work is, appropriately, to present a physical justification as to why the simple mathematical structure we lay out is necessary.
 
@@
 This is reminiscent of a quasistatic approximation.
-Finally, while the approach presented here does not solve the so called ``measurement problem'', it does however give us an understanding of why we have the problem to begin with: our states are really just the equilibrium states of faster scale processes and those cannot, in general, be described by quasistatic ones.
 
-%The connection to measurement is simply that, as we will see later in more detail, such processes are always required if we want to measure a quantity, even in classical mechanics or thermodynamics.
-For example, if we want to measure the temperature of a gas it needs to be in equilibrium for long enough for us to conduct the measurement.
-If we are measuring the position of a ball, we need to integrate over a (small) period of time to integrate the giggling created by the random collisions of air particles.
-That is: if we want to measure the vertical spin component of an electron, something needs to make sure that we actually have and electron and its vertical component is well defined.
-So we need a process whose outputs are equilibria with those properties.
+Finally, we note that the approach presented here does not solve the so called ``measurement problem''\cite{Genovese,Bassi}, at least not in the way it is typically framed.
+We still do not have a description of what happens during an arbitrary process (e.g.
+a projection), yet we have an explanation of why we cannot provide one in terms of a quasistatic evolution (i.e.
+unitary evolution) since a special case cannot describe the general one.\footnote{We purposely leave unspecified whether the faster scale processes are driven by internal or external dynamics, or by a mixture of the two.
+As such, this approach does not technically pick an interpretation: it merely states that whatever happens during a measurement in a given interpretation is also happening at every instant in time.}
 
@@
 For example, as we change the volume of a gas, the region against the moving wall will be compressed sooner than the opposite region, but if we wait a short while we will again obtain a uniform pressure.
-If we want to study that transition, we may conceptually divide the system into small volumes and the pressure is replaced by a pressure field.
-Still, we need to assume each small volume to have enough molecules to be in equilibrium.
-We can go into more detail, and describe the distribution of position and velocities of the molecules.
+If we want to study that transition, we may conceptually divide the system into small volumes and we replace the pressure by a pressure field.
+Still, we need to assume each small volume has enough molecules to be in equilibrium.
+We can go into more detail, and describe the distribution of position and velocity of the molecules.
 Yet, we need to assume that the molecules themselves are not undergoing chemical changes.
 We could study the atoms of the molecules, yet we would have to assume that the atoms are stable under nuclear fission.
-We could study the electrons and nucleons of the atoms, yet we would have to assume they are stable under weak nuclear forces.
-As we move to a finer and finer description, we have faster and faster scale processes that we assume do not change the basic constituents.
+We could study the electrons and nucleons of the atoms, yet we would have to assume they are stable under weak nuclear forces.\footnote{This hierarchy of time scales is indeed used in many areas of science and engineering under the practice of multiscale modeling.
+\cite{Horstemeyer,Weinan} It allows one to break down a complex problem so that the relevant dynamics is studied efficiently at the right level of scale, gaining more insight and leading to faster calculations.
+Each scale will use different models and techniques, such as continuum mechanics, Monte-Carlo methods for statistical physics, molecular mechanics, quantum chemistry and so on.
+\cite{Brandt} During a computation cycle, each level of scale treats the faster and slower dynamics as fixed parameters, which are updated appropriately for the following iteration.} As we move to a finer and finer description, we have faster and faster scale processes that we assume do not change the basic constituents.
 At each level of the description what is assumed to be in equilibrium changes and what the laws describe changes as well.
@@
 States are always equilibria of faster scale processes, which in quantum mechanics are modeled with the projection.
-States are chosen to satisfy a deterministic law of evolution, necessarily describing a quasistatic process, which in quantum mechanics is modeled by the Schroedinger equation.
+States are chosen to satisfy a deterministic law of evolution, necessarily describing a quasistatic process, which in quantum mechanics is modeled by the Schr\"{o}dinger equation.
 This general framework only uses standard physics concepts and puts no special role on the observer except the choice of what processes the system will be subjected to.
 
-Violations of the Schroedinger equation, then, should be expected only for processes that cannot be considered quasistatic.
-If we were to implement a measurement process, then, in a way that the transfer of information between the system and the probe happens in a slow and controlled fashion, this could not provide a violation of the Schroedinger equation, nor would it tell us that all projections are quasistatic processes.
-A better description of the faster scale processes given by the projections, then, will inevitably need a new state space as the standard quantum states only describe equilibria, and we should expect the faster scale processes to go out of equilibrium.
-Ultimately, a description at Plank time scale will fail to be continuous and quasistatic since a finite change in a finite time will not be divided into infinitesimal adiabatic changes over infinitely many time steps.
+Violations of the Schr\"{o}dinger equation should be expected only for processes that cannot be considered quasistatic.
+If we were to implement a measurement process, then, in a way that the transfer of information between the system and the probe happens in a slow and controlled fashion, this could not provide a violation of the Schr\"{o}dinger equation, nor would it tell us that all projections are quasistatic processes.
+On the other hand, a better description of the faster scale processes given by the projections cannot be given in terms of unitary evolution, as this represents the quasistatic case.
+It would inevitably require a new state space as the standard quantum states can only describe equilibria of those processes.
+Moreover, a description at Planck time scale will ultimately fail to be continuous and quasistatic since a finite change in a finite time will not be divisible into infinitesimal adiabatic changes over infinitely many time steps.
 
 To understand the suitable mathematical framework for those regimes where unitary evolution may not hold, we are proceeding by first identifying the physical requirements for all mathematical structures already used in physics, starting from the most basic ones like topological spaces (upon which is based all of differential geometry) and $\sigma$-algebras (upon which is based all of measure theory and probability theory).
-This long term effort is centered mainly around the development of an open access book in which these issues can be explored in full detail and with a rigor more typical of mathematical disciplines.
+This long term effort is centered mainly around the development of an open access book\cite{Carc3} in which these issues can be explored in full detail and with a rigor more typical of mathematical disciplines.
 This rigor forces us to spell out the often hidden assumptions we have while constructing physical theories.
@@
 
-\bibliography{bibliography}
+\begin{thebibliography}{999}
+	\bibitem[Carcassi(2018)]{Carc1}
+	G.
+Carcassi, C.
+A.
+Aidala, D.
+J.
+Baker, L.
+Bieri; From physical assumptions to classical and quantum {Hamiltonian} and {Lagrangian} particle mechanics.
+{\em J.
+Phys.
+Comm.} {\bf 2018}, {\em 2(4)}, 045026.
+
+	\bibitem[Genovese(2010)]{Genovese}
+	M. Genovese; Interpretations of quantum mechanics and measurement problem. {\em Adv. Sci. Lett.} {\bf 2010}, {\em 3}, 249--258.
+
+	\bibitem[Bassi(2013)]{Bassi}
+	A. Bassi, K. Lochan, S. Satin, T. P. Singh, and H. Ulbricht; Models of wave-function collapse, underlying theories, and experimental tests. {\em Rev. Mod. Phys.} {\bf 2013}, {\em 85}, 471.
+
+	\bibitem[Weinan(2011)]{Weinan}
+	E. Weinan; Principles of Multiscale Modeling. {\em Princeton University} {\bf 2011}.
+
+	\bibitem[Horstemeyer(2009)]{Horstemeyer}
+	M.
+F.
+Horstemeyer; Multiscale Modeling: A Review.
+In Leszczyński, Jerzy; Shukla, Manoj K.
+(eds.).
+Practical Aspects of Computational Chemistry: Methods, Concepts and Applications.
+pp.
+87–135.
+ISBN 978-90-481-2687-3.
+
+	\bibitem[Brandt(2002)]{Brandt}
+	Brandt A.
+(2002) Multiscale Scientific Computation: Review 2001.
+In: Barth T.J., Chan T., Haimes R.
+(eds) Multiscale and Multiresolution Methods.
+Lecture Notes in Computational Science and Engineering, vol 20.
+Springer, Berlin, Heidelberg
+
+	\bibitem[Daneri(1966)]{Daneri}
+	A. Daneri, A. Loinger and G. M. Prosperi; Further remarks on the relations between statistical mechanics and quantum theory of measurement. {\em Nuovo Cimento B} {\bf 1966}, {\em 44}, 119--128.
+
+	\bibitem[Merlin(2015)]{Merlin}
+	R. Merlin; A Heuristic Approach to the Quantum Measurement Problem: How to Distinguish Particle Detectors from Ordinary Objects. {\em Int. J. Mod. Phys. B} {\bf 2015}, {\em 29(22)}.
+
+	\bibitem[Tamir(2013)]{Tamir}
+	B, Tamir, E. Cohen; Introduction to Weak Measurements and Weak Values. {\em Quanta} {\bf 2013}, {\em 2}, 7–17.
+
+	\bibitem[Carcassi(2019)]{Carc3}
+	G. Carcassi, C. A. Aidala; Assumptions of Physics. {\em In preparation} {\bf 2019}, http://assumptionsofphysics/book.
+
+\end{thebibliography}
 

### 2019-09-18 Proceedings/2019-Quantum2019/CarcassiQuantum2019Proceedings.tex
Revised directory structure for proceedings

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Trim Size: 9.75in x 6.5in
%% Text Area: 8in (include Runningheads) x 5in
%% ws-ijqi.tex   :   18-6-08
%% Tex file to use with ws-ijqi.cls written in Latex2E.
%% The content, structure, format and layout of this style file is the
%% property of World Scientific Publishing Co. Pte. Ltd.
%% Copyright 1995, 2002 by World Scientific Publishing Co.
%% All rights are reserved.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\newcounter{myctr}
\def\myitem{\refstepcounter{myctr}\bibfont\noindent\ifnum\themyctr>9\else\phantom{0}\fi\hangindent17pt\themyctr.\enskip}
\def\myhead#1{\vskip10pt\noindent{\bibfont #1:}\vskip4pt}

\documentclass{ws-ijqi}
\usepackage{hyperref}
\usepackage[super,sort,compress]{cite}
\begin{document}

%%%%%%%%%%%%%%%%%%%%% Publisher's Area please ignore %%%%%%%%%%%%%%
\catchline{}{}{}{}{}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{THE FUNDAMENTAL CONNECTIONS BETWEEN CLASSICAL HAMILTONIAN MECHANICS, QUANTUM MECHANICS \\ AND INFORMATION ENTROPY}

\author{GABRIELE CARCASSI}

\address{Physics Department, University of Michigan, 450 Church Street\\
Ann Arbor, MI 48109-1040,
United States\\
carcassi@umich.edu}

\author{CHRISTINE A. AIDALA}

\address{Physics Department, University of Michigan, 450 Church Street\\
Ann Arbor, MI 48109-1040,
United States\\
caidala@umich.edu}

\maketitle

\begin{history}
%\received{Day Month Year}
%\revised{Day Month Year}
%\accepted{Day Month Year}
%\comby{(xxxxxxxxxx)}
\end{history}

\begin{abstract}
In a previous work we have shown how the principal difference between classical Hamiltonian mechanics and quantum mechanics is not the physical size of the object being studied, but the degree of accessibility its internal dynamics. Namely, in classical mechanics it is assumed to be fully accessible while in classical mechanics is assumed to be totally inaccessible. As information entropy can be used to characterize how much the state of the whole system characterizes the state of its parts, it provides fundamental insights that allow us to properly understand the analogies and differences between the two theories. Conservation of information, for example, leads to Hamiltonian and Schroedinger evolutions, which describe the deterministic and reversible dynamics. We review
\end{abstract}

\keywords{Information entropy; Hamiltonian mechanics; Quantum mechanics; determinism and reversibility.}

%\tableofcontents  % optional

\markboth{Carcassi, Aidala}
{The Fundamental Connections between Mechanics and Information Entropy}

\section{Introduction}
In a previous work we identified a small set of physical assumptions from which classical and quantum particle mechanics can be rederived. By assumptions here we mean simplifying conditions that our system is supposed to satisfy under the processes under study. Our work differs from others in that the starting points are physical assumptions about the system being studied (e.g whether the parts can be studied or whether the evolution is deterministic) and not some abstract mathematical or information theoretic considerations. It is also different in that both classical and quantum cases are derived on equal footing. That is, we have identified a single point where one assumption leads to the classical case while another lead to the quantum system.

The difference is the answer to a single question: is the internal dynamics of the system accessible or not? If it is fully accessible we recover classical states and if it is totally inaccessible we recover quantum states. This difference can be understood in terms of information entropy and in this paper we develop that insight. We will present here a much shorter version of the classical derivation that carries all the important points. For quantum mechanics, unfortunately we do not currently have one, so we will cover only some of the general ideas. For details on both, we refer to the previous longer paper. We then provide a classical analogue of the quantum uncertainty relationship and see how both can be understood in terms of information entropy conservation under the different initial assumptions.

We believe these insights better clarify the difference between classical and quantum systems, their limit of validity and therefore where one can expect them to fail.

\section{Reducible Systems and Classical Phase Space}
We define a classical system as one that is assumed to be infinitesimally reducible. That is, the system can be divided into parts indefinitely and giving the state of the system is equivalent to giving the state of its parts. For example, we can throw a ball and study its motion; or we can take a red marker, make a dot on the ball and study the motion of the red dot. Studying the motion of the whole ball is equivalent to studying the motion of all the possible red dots we could draw. To a first approximation, a ball can be thought as infinitesimally reducible, made of a continuous material.

If we assume our system is reducible, then the state of the whole system is simply a distribution over the states of the infinitesimal parts, which we call particles. That is, our state is a normalized integrable function
\begin{equation}
\begin{aligned}
	\rho : \mathcal{S} \to \mathbb{R} \\
	\int_{\mathcal{S}} \rho ds = 1
\end{aligned}
\end{equation}
over the particle state space $\mathcal{S}$. If we have a set of state variable $\xi^a : \mathcal{S} \to \mathbb{R}$ that are sufficient to characterize our particles, we can write both the density and the information entropy
\begin{equation}
\begin{aligned}
\rho(\xi^a) &= \rho(s(\xi^a)) \\
I[\rho] &= - \int_{\mathcal{S}} \rho \log (\rho) d\xi^1 d\xi^2 ...
\end{aligned}
\end{equation}
in terms of the state variables.

The issue is that changing coordinates system will change the numeric value of the density. Yet, the density is fully specified by the particle states and, since states are invariant under coordinates transformations, it should be invariant. That is, we have
\begin{equation}
\begin{aligned}
\rho(\hat{\xi}^b) &= \left| \frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho(\xi^a) \\
\rho(s(\hat{\xi}^b)) &= \rho(s(\xi^a)) \\
\end{aligned}
\end{equation}
which means our distribution $\rho$ has to both transform as a scalar (i.e. it fully depends on the state) and as a density (i.e. the integral needs to be invariant).

This apparent contradiction is actually a formal requirement that severely limits what types of manifolds can be used to describe particle states. It turns out that these are precisely the symplectic manifolds: the structure of phase space is exactly the one required to write coordinate invariant distributions. That is, given a set of coordinates $q^i$ (state variables that define the reference frame) we will need another set of state variables $k_i$ with inverse units such that their product will become an invariant pure number. Under a general coordinate transformation $\hat{q}^j = \hat{q}^j (q^i)$ we will simply have 
\begin{equation}
\begin{aligned}
dq^i dk_i = d\hat{q}^j d\hat{k}_j
\end{aligned}
\end{equation}
which means the Jacobian of the transformation are unitary. The invariance of areas is equivalent to the invariance of densities and of information entropy. In other words, the structure of phase space is precisely the one that allows us to define information entropy as coordinate independent. Only on these spaces information entropy can be a physically well-defined quantity over continuous variables.

If we require deterministic and reversible evolution, we require that the density over one state is transported exactly to another state. Equivalently, we require that the information given at one time to identify a part is exactly the same to identify the same part at a different time: information entropy is the same. In this case we are requiring that the time evolution is an area preserving (i.e. canonical) transformation. This gives us Hamilton's equations. This tells us that the following four concepts
\begin{itemize}
	\item Deterministic and reversible evolution - state densities are mapped one-to-one
	\item System isolation - the evolution of the system depends only on its state and on nothing else
	\item Conservation of information entropy - the information required to describe the system does not change in time
	\item Conservation of energy - Hamiltonian evolution
\end{itemize}
are the same concept from different angle.

\section{Irreducible systems and quantum mechanics}

Conversely, we define a quantum system as one that is assumed to be irreducible. That is, the state of the system tells us nothing about the state of the parts. For example, we can isolate an electron and study its motion; yet we cannot put a red mark on part of the electron, we cannot scatter a photon off of only a part of an electron. As the interaction is always with the whole electron, no information can be gained about its internal dynamics. An electron can be thought as irreducible.

We stress that both these assumptions are operational and depend not only on the system, but also on the processes we are studying or available to us. For example, a proton can be modeled as a single quantum system and will behave as such provided the internal dynamics is not relevant. It will exhibit diffraction and interference patterns, have half-integer spin and so on. If we interact in a way that disturbs the internal dynamic, however, we will see that there is internal dynamics in terms of quarks and gluon. In those regimes the proton cannot be treated a single quantum system, but as a composite system.

We weren't able to prepare a shortened version of the quantum derivation, but the gist is that, since the state of the system does not capture the state of the parts, we can imagine these continuously permutating without affecting the dynamic of the whole. The internal motion is the characterized by random variables, a pair $(A,B)$ for each degree of freedom, which maximize the entropy of the distribution of the parts. This means that the variable along each degree of freedom are independent and only the second moments are significant (i.e. statistically equivalent to a multivariate Gaussian). A complex number $\sigma_A + \imath \sigma_B$, then, can be used to keep track of a variable within a degree of freedom and ultimately leads to the state space of quantum mechanics. The norm of the quantum number represents the strength of the internal random process, which is proportional to the size of the system, and the phase is the correlation angle. Superposition, then, means superimposing the random motions which will interfere constructively where correlated and interfere destructively where anti-correlated.

We want to stress, again, that we consider the classical state to be a matter distribution $\rho(q^i, k_i)$. Therefore, when comparing classical and quantum mechanics, we should be comparing the properties of our wave function $\psi(q^i)$ to the matter distribution $\rho(q^i, k_i)$, not to points of phase space. A point-particle is the limit where the extent of the distribution can be considered small, and therefore the center of mass is enough for our description. This limit can be performed within both classical and quantum mechanics if the scale of the problem allows, and, because of Ehrenfest theorem, the behavior is the same. In light of the assumptions, it should be obvious this should be the case: the difference between classical and quantum is in the description of the parts; if the length scale are much greater than the system, then it doesn't matter what we assume about the parts.

\section{Uncertainty principles and information entropy}

Since classical Hamiltonian evolution conserves information entropy, the space of distribution explored by time evolution is limited by that constrain. We then ask the following question: what is the distribution that, given a certain amount of information entropy, minimizes the spread over phase space? That is, what is $\rho(q, k)$ such that $\sigma_q\sigma_k$ is minimized? Given the constrain, no Hamiltonian evolution can lead to a state less spread than that.

It turns out this distribution is the product of two Gaussian distributions
\begin{equation}
\rho(q,k) = \frac{1}{2\pi\sigma_q\sigma_k} e^{- \frac{(q - \mu_q)^2}{2 \sigma_q^2}} e^{- \frac{(p - \mu_p)^2}{2 \sigma_p^2}}
\end{equation}
with information entropy
\begin{equation}
I_G = 2 \pi e \sigma_q \sigma_k.
\end{equation}
Since this is the distribution that minimizes the spread, during the evolution we have
\begin{equation}
\sigma_q \sigma_k \leq \frac{I_0}{2 \pi e}
\end{equation}
where $I_0$ is the initial information entropy. This classical uncertainty relationship is reminiscent of the quantum uncertainty relationship for both the form and the fact that the equality holds for the Gaussian packages. The analogy can be made stronger with a few extra conditions, though we will not expand that here. What we want to show here is the connection between this uncertainty relationship, information entropy, deterministic and reversible evolution and the difference in reducibility between a classical and a quantum system.

First of all we note that this classical uncertainty relationship is given by conservation of entropy and therefore deterministic and reversible evolution. We are free to set the initial amount of information entropy for the classical distribution. Once it is set, this cannot be changed under Hamiltonian evolution. Quantum mechanics has a stronger condition: it is not possible to create a state that violates the uncertainty. That is, we do not have the initial freedom we have in classical mechanics. For quantum states, all pure states have the same information entropy which is set to zero. This difference is a direct consequence of the difference in reducibility: in classical mechanics we can always access the internal dynamics, and therefore we can create narrower and narrower distributions (e.g. we can always divide a continuous material) and study their individual evolution; in quantum mechanics we cannot (e.g. we cannot prepare half an electron) so all states necessarily are associated with the same information entropy. The entropy for a pure state, then, is set to zero to signify that the dynamics within a quantum system is not accessible, there is no information to be had.

The deterministic and reversible case in quantum mechanics is given by unitary evolution, which in fact conserves Von Neumann entropy because each pure state is mapped to a pure state. Conversely, the Lindblad equation allows the entropy to increase and corresponds to irreversible evolution. However, in that case it is the entropy of the mixed case that is being increased. In classical mechanics this would correspond to mapping one initial distribution to multiple final distribution with the same entropy according to different probability. That is, we have a probability distribution over matter distributions. As each matter distribution has the same entropy, it is the spread in probability that increases the entropy. But in classical mechanics we can also do something more: we can have a density distribution spread more or less over phase space with a dynamic that is not Hamiltonian and change the information entropy of the matter distribution itself. For example, a damped harmonic oscillator would concentrate the distribution around the equilibrium and decrease the entropy. This type of dynamic cannnot be described in the context of quantum mechanics. Another aspect of this same issue is that given two quantum states we can always find a unitary evolution that connects the two (at least mathematically) while this is not possible in classical Hamiltonian mechanics (e.g. two distributions with different entropy cannot be linked by Hamiltonian evolution).

We believe these insights clarify two issues. The first is that the difference between a classical and quantum system is not primarily about is size. The previous discussion does not depend on the physical extension of the system or about the number of its degrees of freedom. It is about the accessibility of the internal dynamics. If we are to look at scenarios where classical mechanics fails, we simply have to look at cases where we can no longer assume that the internal dynamics is indefinitely accessible. In a Bose-Einstein condensate, in a pair of entangled photons or in a diffracting neutron the parts cannot be studied independently without significantly altering the system: infinitesimal reducibility does not apply. Conversely, if we are to look at scenarios where quantum mechanics fails as applied for a specific system, we simply have to look at cases where the internal dynamic of that system is indeed accessible.

The second issue is that it tells us what is being quantized to begin with. Quantities like energy, spin or momentum are indeed quantized in some cases, but they are not always quantized in the same way and there is a continuum of different cases. For example, spin along one direction is quantized, but we can continuously change the direction. Energy is quantized only for bound systems, and even in those cases a relative boost can continuously change the value. The entropy associated with all pure states, instead, is always the same for all states in all frames. Again, this is not true in classical mechanics: a matter distribution over phase space can have an arbitrary information entropy. The entropy associated to a pure quantum state, which is itself a distribution (i.e. a wave function), is always the same. That is what is quantized. It if weren't, the Von Neuman expression would not be the correct expression for entropy as each pure state would have to be weighted different. The lack of that term means all pure states carry the same contribution: the information entropy of pure states is what is quantized.

\section{Submitted abstract}
In a recent paper [1] we have shown that classical and quantum Hamiltonian particle mechanics can  be  rederived  from  a  handful  of  physical  assumptions.  We  will  present  and  expand  a consequence   of   that   work,   namely   that   classical   Hamiltonian   mechanics   coincides   with conservation of information entropy, and how this insight allows us to better understand quantum mechanics as well.We  will  start  by  showing  that  classical  Hamiltonian  mechanics  can  be  recovered  from  two fundamental assumptions. The first is that the system we study is infinitesimally reducible: it is made  of  infinitesimal  parts  and  the  state  of  those  parts  is  equivalent  to  the  state  of  the  whole system. The state of the overall system is then a distribution over the states of the infinitesimal parts. As states must be defined independently of coordinate system (i.e. of observer), the density over the states of the parts must be invariant under coordinate transformation. This allows us to recover  the  geometrical  structure  of  phase  space  (i.e.  state  variables  organized  in  conjugate pairs). The second assumption is that of deterministic and reversible evolution: knowing the state at one time is enough to predict the state at future times and reconstruct the state at past times. This means that the density for a given state will be mapped exactly to a future state, which will give us Hamiltonian evolution. We then show that requiring densities to be coordinate invariant and  not  change  in  time  is  equivalent  to  requiring  that  the  information  entropy  (i.e.  the  number bits of identify an element of the distribution up to unit precision) is conserved.Throughout the derivation, we will see how classical analogues of quantum mechanics (i.e. wave number, uncertainty principle, particle/anti-particle states) naturally emerge. We will discuss how quantum mechanics differs from classical mechanics by assuming irreducibility instead: that the system is still made of infinitesimal parts but that the state of the overall system tells us nothing about  the  state  of  the  parts.  We  will  see  how  this  qualitatively  leads  to  the  known  quantum effects.This work is part of a larger project [2] that aims to rederive the known laws of physics from a handful of physical assumptions.

\section{Conclusion}

We have shown that the principal difference between classical and quantum systems is how much the state of the whole can say about it parts. This can be understood in terms of information entropy: classical states as distribution over phase space allow for ever decreasing information entropy while quantum states has wave functions have a fixed entropy which is set to zero. Deterministic and reversible evolution conserve information, therefore we have an uncertainty relationship under Hamiltonian evolution.

We conclude by noting that both the classical assumptions (i.e. the internal dynamic is infinitesimally accessible) and the quantum assumption (i.e. the internal dynamic is totally inaccessible) seem untenable. If we were to suppose that the internal dynamics, though not influenceable, had consequences (e.g. they would decide when a particle decays or how an entangled system is broken), then we would also have processes that increased entropy over time (i.e. more information is needed to identify future state). Such a change of assumption, though, is not trivial as it would necessarily require rethinking the state space as well.

\section*{Acknowledgments}

This section should come before the Appendices and References. Funding
information may also be included here.

\renewcommand\bibname{References}

\bibliographystyle{ws-ijqi}
\bibliography{bibliography}

\end{document}

### 2019-09-18 Proceedings/2019-Quantum2019/CarcassiQuantum2019Proceedings.tex
Bibliography now works

@@
 \usepackage[super,sort,compress]{cite}
+\usepackage{url}
 \begin{document}
@@
 \section{Introduction}
-In a previous work we identified a small set of physical assumptions from which classical and quantum particle mechanics can be rederived.
+In a previous work\cite{Carc1} we identified a small set of physical assumptions from which classical and quantum particle mechanics can be rederived.
 By assumptions here we mean simplifying conditions that our system is supposed to satisfy under the processes under study.

### 2019-09-25 Proceedings/2019-Quantum2019/CarcassiQuantum2019Proceedings.tex
Adding citation

@@
 By assumptions here we mean simplifying conditions that our system is supposed to satisfy under the processes under study.
-Our work differs from others in that the starting points are physical assumptions about the system being studied (e.g whether the parts can be studied or whether the evolution is deterministic) and not some abstract mathematical or information theoretic considerations.
+Our work differs from others\cite{PhysRevA.84.012311} in that the starting points are physical assumptions about the system being studied (e.g whether the parts can be studied or whether the evolution is deterministic) and not some abstract mathematical or information theoretic considerations.
 It is also different in that both classical and quantum cases are derived on equal footing.

### 2019-09-26 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Incorporating comments from Julian and reviewing

@@
 
-\title{Universal Characterization of Variability by \\Shannon's $\sum_ip_i\,\textrm{ln}\,p_i$ Formula}
+\title{Universal Characterization of Physical Variability by \\Shannon's $-\sum_ip_i\,\textrm{ln}\,p_i$ Formula}
 
 \author{
-	Julian Barbour, 
+	Julian Barbour and 
 	Gabriele Carcassi
@@
 One of us recalls but has been unable to relocate a sentence in one of Leibniz's many writings to the effect that reality \emph{is} variety.
-This is consistent with his view that any contingent thing is defined by a collection of attributes, which necessarily must be distinguishable
+Whether he made the statement or not, it is consistent with his view that any contingent thing is defined by a collection of attributes, which necessarily must be distinguishable
 at some epistemic level.
@@
 
-Without committing ourselves to Leibniz's overarching principle, our aim in this paper is to draw attention to the specific nature of the variety that can be observed in the universe and the unique suitability of Shannon's entropy (or information) formula $\sum_ip_i\,\textrm{ln}\,p_i$ for the characterization of the variety.
+Without committing ourselves to Leibniz's overarching principle, our aim in this paper is to draw attention to the specific nature of the variety that can be observed in the universe and the unique suitability of Shannon's entropy (or information) formula $-\sum_ip_i\,\textrm{ln}\,p_i$ for the characterization of the variety.
 Indeed, as we will show in the next section, opportunities for application of Shannon's formula are ubiquitous both on the earth and in the universe at large.
+PRECISELY BECAUSE OF THIS GENERALITY WE WILL OFTEN USE THE EXPRESSION SHANNON `FORMULA' RATHER THAN `ENTROPY' OR `INFORMATION'.
 
-We also wish to comment on the various interpretations of Shannon's formula and attempt to answer this question: what does the formula represent?
+We also wish to comment on the various interpretations and uses of Shannon's formula and attempt to answer this question: what does the formula represent?
 It is well known that when Shannon recognized the importance of his formula in communication theory he sought von Neumann's advice on the name he should give to the expression he had found.
@@
 More examples will follow in the next section.
+In this paper we do not ask why the universe sustains the variety that it does.
+We take it as a given.
 
-\section{Comments on Boltzmann's $H$ Theorem}
+NEW PARAGRAPH
 
-[This is for you Gabriele].
-In Boltzmann's 1872 paper, he introduced the function that later was called the $H$ function, though it seems he meant to call it [upper case] $\eta$ and it was printed as $E$ (which somehow later became $H$).
-In the paper Boltzmann describes (in $\mu$-space) a gas of monatomic molecules that has already become spatially uniform and also has the same distribution wrt the particles' kinetic energies $x$ at each space point.
-He therefore considers a spatially constant distribution function $f(x,t)$ and wants to show that whatever its initial value it will tend in accordance with the Boltzmann equation (which he has previously derived and I won't give here) to the Maxwell distribution $f(x,t)=C\sqrt xe^{-hx}$ as $t\rightarrow\infty$.
-He defines his function $E$ as follows:
-\begin{equation}
-E=\int_0^\infty f(x,t)\left\{\textrm{log}\left [{f(x,t)\over \sqrt x}\right ]-1\right\}\textrm dx\label{hfun}
-\end{equation}
-and aims to prove that it cannot increase and attains its least value when $f(x,t)$ becomes the Maxwell distribution.
-Boltzmann does not give the argument that leads him to the function (\ref{hfun}), but somewhat later in the paper he notes that the $-1$ in the integrand merely reflects the fact that
-$$
-\int_0^\infty f(x,t)\textrm dx
-$$
-is the total number of molecules in unit volume, so that 
-\begin{equation}
-E=\int_0^\infty f(x,t)\textrm{log}\left [{f(x,t)\over \sqrt x}\right ]\textrm dx\label{hfunb}
-\end{equation}
-also cannot increase.
+We conclude this section with a word on terminology.
+In mathematics, Cantor defined a set as ``a collection into a whole $M$ of definite and separate objects''.
+He called the objects \emph{elements}.
+We will also use the word `elements' to denote the things we consider but subject to the proviso that they have a distinguishing feature (or features) that allows us to say that \emph{they are of the same kind}.
+Thus, we consider sets of a restricted type.
+The point is this.
+To constitute a set in accordance with Cantor's definition, it is only necessary for its elements to be \emph{definite and separate} objects.
+This leaves total freedom as to how such objects in the universe are to be parcelled up into specific sets.
+For example, a set might consist of Mars and the word `go'.
+We think such latitude misses one of the most significant aspects of the universe, namely, that the things within it can be distinguished by attributes that are much more specific than the two required by Cantor.
+Our field of investigation goes beyond mere number, which washes out all attributes apart from the two, of being definite and separate, that allow us to count.
+They were the only attributes that Cantor needed; we need more.
 
-Boltzmann does not say how he arrived at the function (\ref{hfun}), but its form is obviously dictated by the form of the Maxwell distribution.
-I'm terrible at equations, but I assume the $\sqrt x$ in the denominator of the integrand in (\ref{hfunb}) is because the usual Shannon expression is `skewed' by the subsidiary Lagrangian multiplier that fixes the total energy to a definite value.
-Later in the paper, when considering polyatomic molecules, he arrives by means of arguments I have not yet attempted to understand at the familiar $f\, \textrm{log}\,f$ form.
-
-In his very concise 1860 derivation of his energy distribution function, Maxwell commented that
-
-\begin{quote}\small
-
-It appears from this proposition that the velocities are distributed among the particles according  to the same law as the errors are distributed among the observations in the theory of the ``method of least squares''.
-
-\end{quote}\normalsize
-
-In his much more substantial 1867 paper, Maxwell noted that his 1860 derivation could be seen as `precarious' and gave an alternative proof.
-However, this does not eliminate the mathematical identity of his distribution with the error distribution.
-Would you agree that this (somewhat indirectly) establishes another example of variablity within a distribution of things [which I think we might use instead of `elements'; `members' (of a set) is another word we might consider]?
-
-\vspace{2in}
+\vspace{.5in}
 
@@
 We have a set of elements $E = \{e_\alpha\}_{\alpha=1}^N$.
-These can represent the galaxies in the Laniakea Supercluster, the animals in the Gal\'{a}pagos islands, the molecules in a box of gas given a particular macrostate, the different takes of spin measurement for a given quantum state, or the words in the complete works of Shakespeare.
+In accordance with the comments made in our introduction, they can for example represent the galaxies in the Laniakea Supercluster, the animals in the Gal\'{a}pagos islands, the molecules in a box of gas given a particular macrostate, the different outcomes of spin measurement for a given quantum state, or the words in the complete works of Shakespeare.
 The choice of the particular set is driven by the interest and practices of scientists and their fields of study and therefore it is, in this sense, discretional.
@@
 These can represent the galaxy types, the genus of the animal, the possible states of the molecules, the possible values for spin or the spelling of words.
-This will give us a sequence $\{q(e_\alpha)\}_{\alpha=1}^N$ of the descriptions that are associates with each element in the set.
+This will give us a sequence $\{q(e_\alpha)\}_{\alpha=1}^N$ of the descriptions that are associated with each element in the set.
 Again, the choice of properties is ultimately discretional, grounded on what particular aspects of the elements we are studying in a given case.
@@
 
-Now that we have constructed our distribution, we want to construct an indicator $S$ that quantifies how much variability the elements exhibit within the distributions.
+Once we have constructed our distribution, we will want to construct an indicator $S$ that quantifies how much variability the elements exhibit within the distributions.
 That is, we want to quantify the degree of diversity that the values can have within the distribution.
@@
 We may be tempted to use standard statistical quantities, like the range or the variance, but this is not possible.
-First, if the values associated to the bins are non-numerical (e.g.
+First, if the values associated with the bins are non-numerical (e.g.
 types of galaxies, words), statistical quantities are not well defined.
-Second, permutation of the values does not have an impact on variability of the elements, while statistical quantities will in general be affected.
+Second, relabeling the values (e.g.
+switching names, switching units, non-linearly changing coordinates) does not have an impact on variability of the elements, while statistical quantities will in general be affected.
 This tells us that our indicator cannot depend on $Q$ itself, but only on the weights $p_i$.
-That is, we require that $H=H(p_i)$.\footnote{Note that for continuous quantities the weights $p_i$ are affected by the choice of $Q$.
-We will concentrate on this problem later.}
+That is, we require $H=H(p_i)$.\footnote{Note that for continuous quantities the weights are densities and are affected by the choice of $Q$: the unit is required to specify the numeric value (e.g.
+$1 \% / m$) and this will change under unit transformation (e.g.
+$1000 \% / Km$).
+Therefore, as the weights $p_i$ themselves depend on the unit, the indicator $H(p_i)$ will in general depend on the choice of $Q$.
+Though this is source of additional confusion, for the purpose of defining a measure of variability it does not change things conceptually.
+We will discuss the problem in connection to entropies in thermodynamics and statistical mechanics, since coordinate invariance is paramount in physics.}
 
-We  expect small changes in a distribution, small changes of the weights, to produce small changes in variability, therefore we require $H$ to be a continuous function with respect to the $p_i$.
-We also expect that if we have a uniform distribution $p_i = 1/I$, then greater $I$ should correspond to greater variability.
-Therefore, in the case of uniform distribution, we require $H$ to be monotonically increasing with $I$.
+We  expect small changes in a distribution -- small changes of the weights -- to produce small changes in variability; we therefore require $H$ to be a continuous function of the $p_i$.
+We also expect that as the number of values found within the distribution increases, so will the variability.
+Therefore if we have a uniform distribution over $I$ cases, so that $p_i = 1/I$, we require $H$ to be monotonically increasing with $I$.
 
@@
 \item If $p_i=1/I$ then $H$ is a monotonically increasing function of $I$
-    \item Let $p_i$ and $q_j$ be the weights for two distribution respectively over $I$ and $J$ bins.
+\item Let $p_i$ and $q_j$ be the weights for two distributions respectively over $I$ and $J$ bins.
 Let $r_k$ be the distribution over $K=I+J-1$ bins constructed by expanding the $a$-th bin of the first distribution using the second distribution.
@@
 The Shannon entropy, then, is a clear and important attribute of a distribution: it quantifies the variability of the elements within the distribution.
-It quantifies the variety of values within the distribution.
+It quantifies the variety of values one finds within the distribution.
 It is true that the choice of the elements and binning are not fixed, it is true that the meaning of the weights depends on what the distribution is describing, and therefore that the precise meaning of this variability is context dependent.
 But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope a probability and so on.
-And like a real number represents a quantity, transcending what is physically quantified, the Shannon entropy represents the variability of the elements in a distribution, transcending what the distribution physically signifies.
+And like a real number represents a quantity, transcending what is physically quantified, the Shannon entropy represents the variability of the elements\footnote{We again stress that out elements are physical and therefore posses some property beyond the bare `definite and separate' of Cantor's definition, which was perfectly suited to study numbers.
+Without these extra attributes, which are `washed out' by said definition, we would not be able to create a distribution in the first place.} in a distribution, transcending what the distribution physically signifies.
 
@@
 \begin{enumerate}
-    \item There is 50\% chance you won one million dollars in the lottery and 50\% chance you didn't;
+\item There is 50\% chance you won one million dollars in the lottery and 50\% chance you didn't.
 \item There is 50\% chance you won one million dollars in the lottery and 50\% chance you won half a million dollars.
@@
 
-Unfortunately, entropy in general is often associated with vague characterizations as the ones presented.
-It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending by the authors, which leads to confusion and misunderstanding.
-The characterization of Shannon entropy we have given, on the other hand, applies to all cases, it is enough to understand the assumptions required to rederive it and therefore is more fundamental.
+Unfortunately, entropy in general is often associated with vague characterizations like the the two presented.
+It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending by the authors[CITATION NEEDED], which leads to confusion and misunderstanding.
+The characterization of the Shannon variability measure we have given, on the other hand, applies to all cases and leads naturally to the assumptions required to rederive it.
+OUR CHARACTERIZATION IS therefore is more fundamental.
 If we look at the galaxies in our universe, what variability do they exhibit in terms of their types?
@@
 If we look at the molecules in a given macrostate, what variability do they express in terms of their microstates?
-If we look at the words in the complete works of Shakespeare, what variability we find in its vocabulary?
-All these questions are interesting, all these questions are profound and all these questions are answered by a single concept: the Shannon entropy.
+If we look at the words in the complete works of Shakespeare, what variability we find in his vocabulary?
+All these questions are interesting, all these questions are profound and the answer to all these questions are expressed through just one thing: the Shannon variability measure.
 
-If the Shannon entropy is a measure of variability, why is it connected to information? What is it measuring and in what units?
+If the Shannon expression is a measure of variability, why is it connected to information? What is it measuring and in what units?
 
@@
 We can gain insight with the following example.
-Suppose we fix a distribution, say the animals in the Galap\'{a}gos islands.
-Suppose I pick a specific animal from the distribution and you want to know its species.
-Suppose the only way for you to get that information is to ask me a series of questions.
-Suppose each questions has two possible answers, yes or no.
+Suppose we fix a distribution, say the animals in the Galap\'{a}gos islands grouped by their respective species.
+Suppose we pick a specific animal from the set and you want to know its species.
+Suppose the only way for you to get that information is to ask a series of questions with only two possible answers, yes or no.
 How many questions would you have to ask?
@@
 Others, like ``is it an American Flamingo (Phoenicopterus ruber)?", would give us a lot of information in the positive case but little in the negative case.
-It should be clear, though, that there must be a limit to how few questions you must ask to get to the answer: a single yes/no question cannot be enough.
-It should, hopefully, be intuitive that to a greater variability within the distribution will correspond a greater number of questions you must ask.
-That is exactly what the Shannon entropy measures: the minimum average number of questions one has to ask to identify a value in the distribution.
+However, it is clear that there is always a minimum number of questions that must be asked to get to the answer.
+Given that there are more than two species, a single question cannot be enough.
+It should however be intuitively clear that to a greater variability within the distribution will correspond a greater number of questions you must ask.
+That is exactly what the Shannon variability quantifies: the minimum average number of questions one has to ask to identify a value in the distribution.
 It gives us the number of questions for an ideal strategy for our game of Twenty Questions.
@@
 That is why some people say the entropy is ``lack of information", because it's the information the distribution does not have.
-But, again, this is just missing the point of what it is being described.
+But, again, this is just missing the point of what is being described.
 If one is not interested in identifying elements there is no ``missing information".
@@
 
-As we saw in the previous sections, Shannon entropy is a very general concept that applies to any type of distribution, no matter what the problem domain.
-To understand its link to the thermodynamic/statistical mechanics entropies we need two answer two questions.
-First, what does the Shannon entropy represents in the context of statistical mechanics and why is it a state variable?
-Second, why does the Shannon entropy have the same properties of the thermodynamic entropy?
-That is, we have to understand why it increases during irreversible evolution, it is conserved during reversible evolution, it is maximized at equilibrium and it is additive for independent systems.
+As we saw in the previous sections, Shannon variability is a very general concept that applies to any type of distribution, no matter what the problem domain.
+It does not require to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamic and statistical mechanics which requires them.\footnote{For this reason it is unfortunate that it was given the same name.} It is only when variability is appropriately applied to this particular setting that, though not apparent at first, a connection can be made.
+To understand this link we need to answer two basic questions.
+First, what does the Shannon variability represent in the context of statistical mechanics and thermodynamics?
+In particular, why is it a state variable and what property of the system is it describing?
+Second, why does the Shannon variability satisfy the same properties as the entropy?
+Namely, why does it increase during irreversible evolution, why is it maximized at equilibrium and why is it additive for independent systems?
 
-The first question is straight-forward to answer.
+The first question is straightforward to answer.
 The macroscopic variables of a system (e.g.
-volume, pression, energy) do not give a full description of its pars.
-If we fix a macrostate and then measured the state of all microscopic constituents, we would not find the same answer every time.
+volume, pressure, energy) do not give a full description of its parts.
+If we fix a macrostate and then measure the state of all its microscopic constituents, we will not get the same answer every time.
+This true in the case of non equilibrium (e.g.
+we suddenly compress a gas pushing a piston) and of equilibrium (e.g.
+we give enough time to equilibrate).
 The choice of a macrostate, then, identifies a set of possible microstates that are compatible with it.
-It is then valid to ask the question: what is the variability of microstates one can find within a given macrostates?
-This would be answered by the Shannon entropy, which will measure the variability in bits required to identify a microstate given the macrostate.
-This quantity is a property of the macrostate: it is a state variable.
+It is then valid to ask the question: what is the variability of microstates one can find within a given macrostate?
+This is answered by the Shannon variability, which gives us the number of bits required to identify a microstate given the macrostate.
+Since the macrostates fully defines the distribution of possible microstates, this distribution does not depend on the particular way the macrostate came to pass.
+It is path independent: it is a state variable.
+In short, as a macrostate is a distribution over microstates, it will admit as a state variable the Shannon variability, regardless of whether we know the actual distribution.
 
-For the others, suppose we have a volume of gas at a particular temperature and pressure in perfect equilibrium.
-While the microstate is constantly changing, the macrostate remains the same.
-Effectively, the equilibrium is simply permutating elements of the distribution.
-As we saw, permutations do not change the variability which has to remain constant.
-At equilibrium the Shannon entropy has to remain constant.
+Note, though, that the thermodyanmic entropy is defined differently.
+It is defined over a succession of equilibrium states by noting that the integral of the heat exchanged divided by the temperature is path independent.
+One then chooses a particular state $A$, assigns it an arbitrary value $S_A$ and the entropy for all other states is given by the integral $S(B) = S_A + \int_A^B \frac{dQ}{T}$.
+At this point, it is not apparent why these two quantities defined independently would be related.
+What we can say, though, is that all quantities that increase over irreversible processes, are maximized at equilibrium and that are additive for indipendent systems are precisely the quantities that are linear functions of entropy.[CITE AXIOMATIC WORK] A linear function of entropy, though, simply means a redefinition of $S_A$ and a change of units for either $dQ$ or $T$.
+Therefore if we can show that the Shannon variability possesses those three properties than it is equivalent to the entropy up to a change in units, which brings us to the second question.
 
-Suppose we compress the gas adiabatically and then let it re-expand adiabatically to the original state.
-As the system is isolated, the whole process is deterministic and reversible.
-An initial microstate has to map with a specific intermidiate and final microstate.
-Identifying an element at each stage is equivalent to identifying an element in any state.
-The variability within the distribution at each step, then, cannot change: the same number of bits, the same information, must be enough to identify initial, intermediate or finale microstates.
-Deterministic and reversible processes conserve Shannon entropy.
+Suppose we have an irreversible process, which means we can start from many out-of-equilibrium states which end into a single equilibrium state.
+During the whole process, some of the macro variables are going to stay constant as they characterize the constraint the system is subjected during the evolution.
+For example, if the walls are fixed, the gas will have a set volume; if the walls are insulated, then the average energy will be conserved.\footnote{It can also happen that the irreversible process does have a single possible equilibrium, as in the case of spontaneous symmetry breaking.
+New variables will identify the equilibrium which we can think of new constraints appeared once the initially symmetry is broken.} As different out-of-equilibrium macrostates merge during the evolution, they merge the possible initial conditions that lead to the same macrostate.
+The variability of microstates within the initial out-of-equilibrium macrostate, then, must be smaller than the variability of the microstates within the final equilibrium state.
+The amount of information needed to identify a microstate at equilibrium allows us to identify the initial microstate across multiple out-of-equilibrium states and therefore is greater than the information needed to identity a microstate in an out-of-equilibrium macrostate.
+Irreversible processes increase the Shannon variability within the macrostate, which corresponds to a greater uniformity of the macrostate itself since uniform distributions are the ones that maximize the variability.
 
-Now suppose we have a process that starts from an out-of-equilibrium state and ends in an equilibrium.
-The initial state will not be fully described by the standard macroscopic variables.
-In fact, there will be many out-of-equilibrium states that will lead to the same final equilibrium.
-In this sense the process is irreversible because knowing the final macroscopic variables will not be enough to reconstruct the initial macroscopic state.
-Different possible initial out-of-equilibrium states merge into a final equilibrium state.
+To understand why this is maximized at equilibrium, suppose it weren't.
+That is, suppose there was another macrostate which satisfied the same constraints, but for which the Shannon variability was greater.
+If we prepared such a state and let it equilibrate, it could not evolve to the equilibrium because that has lower Shannon variability.
+Its equilibrium, then, would be different from the previous one, but this cannot be since we said that those constraints would lead to a single equilibrium state.
+Therefore the Shannon entropy is maximized for an equilibrium state.
 
-If we assume, though, that the microstate evolves deterministically and reversibly, knowing the final microstate would allow us to reconstruct the initial microstate and what initial out-of-equilibrium macrostate would be compatible with it.
-As different out-of-equilibrium macrostate merge during the evolution, they merge the possible initial conditions that lead to that same macrostate.
-The variability of microstates within the initial out-of-equilibrium macrostate, then, must be smaller than the variability of the microstates withing the final equilibrium state.
-The amount of information needed to identify a microstate at equilibrium allows us to identify the initial microstate across multiple out-of-equilibrium states and therefore is greater than the information needed to identity a microstate in an out-of-equilibrium macrostate.
-Irreversible processes increase the Shannon entropy which, as we saw before, stabilizes at equilibrium.
+Now suppose we have two independent systems, like two volumes of gasses.
+As they are independent, picking the microstate of one system does not constrain the microstate of the other.
+That is, the distribution over microstates are statistically independent.
+This means that the statistical distribution is the product of the join distributions and, in this case, the Shannon entropy is additive.
 
-Now suppose we have two independent volume of gasses.
-This means that picking the microstate of one system does not constrain the microstate of the other.
-That is, the distribution over microstates are statistically independent.
-In this case, the Shannon entropy is additive.
-
-The Shannon entropy, then, has all the right properties to establish a conceptually meaningful connection with the thermodynamic/statistical mechanics entropies.
-Yet, there are some mathematical details that need to be worked out.
+The Shannon variability, then, has all the right properties to establish a conceptually meaningful connection with the thermodynamic/statistical mechanics entropy.
+It is indeed an independent concept but when applied to macrostates yields a state variable with the right conceptual characteristics.
+There are still some mathematical details that need to be worked out to have a complete picture.
 These details don't change the overall intuitive picture, but they present yet again opportunities for confusion.
 
-The first is the role of the Boltzmann constant $k_B$, which is a conversion factor between the Shannon entropy $S = - \sum p_i \log(p_i)$ and the Gibbs entropy $S_G = - k_B p_i \sum \log(p_i)$.
-The issue here is that temperature, internal energy and entropy are typically linked by the relationship $T = \frac{\partial U}{\partial S}$.
+The first is the role of the Boltzmann constant $k_B$, which is a conversion factor between the Shannon variability.
+$S = - \sum p_i \log(p_i)$ and the Gibbs entropy $S_G = - k_B p_i \sum \log(p_i)$.
+This conversion constant should not surprise us, since we only established a linear relationship between variability and entropy.
+The question is whether this conversion itself is has a physical meaning, or it is also simply due to a choice of units.
+The standard relationship between temperature, internal energy and entropy is $T = \frac{\partial U}{\partial S}$.
 Historically, temperature and internal energy are the quantities that were defined first.
@@
 In this setting, $U$ is measured in Joules, $S$ in nats (which are dimensionless), $\beta$ is measured in inverse Joules.
-Temperature is still measured in Kelvin, but we do not regard it as a primary objects.
-We find this approach, sometimes called entropy first, prefereable for a number of reasons, expecially in statistical mechanics.
+Temperature is still measured in Kelvin, but we do not regard it as a primary object.
+We find this approach, sometimes called entropy first, preferable for a number of reasons, especially in statistical mechanics.
 First, entropy is invariant under coordinate transformations and therefore it is better to keep it a pure number and not mix it with units of energy, which is not invariant under boost.\footnote{This is essential if one ever hopes to put statistical mechanics in a relativistic setting.} Second, the maximization of entropy at equilibrium guaranties us that $S(U)$ is a single valued function (i.e.
 the maximum given the value of $U$) and $\frac{\partial S}{\partial U}$ is therefore well defined.
-On the other hand, $U(S)$ is not in general single value (e.g.
+On the other hand, $U(S)$ is not in general single valued (e.g.
 in the Ising model).
 This means that temperature can be infinite where the curve inverts, where small changes of energy do not change the entropy, and can be negative in the other region.
-That is, the infinitely cold system is temperature zero, which cannot be achieved, as it warms up temperature rises until it become infinite, and as the system becomes even hotter the temperature is negative and increase toward zero, which would correspond to an infinitely hot system.
+That is, the infinitely cold system is temperature zero, which cannot be achieved, as it warms up temperature rises until it becomes infinite, and as the system becomes even hotter the temperature is negative and increase toward zero, which would correspond to an infinitely hot system.
 On the other hand, $\beta$ is always finite and cannot be infinite.
-An infinitely cold system, then would correspond to positive infinite $\beta$, as the system warms up beta decreases, passes from zero, becomes negative and an infinitely hot system corresponds to minus infinite $\beta$: this is a much better property for a physically meaningful quantity.
+An infinitely cold system, then would correspond to positive infinite $\beta$, as the system warms up beta decreases, passes from zero, becomes negative and an infinitely hot system corresponds to minus infinite $\beta$: this is a much better behavior for a physically meaningful quantity.
 Third, when setting up entropy maximization, $\beta$ is quite naturally the Lagrange multiplier associated with energy conservation, therefore mathematically it has a much clearer role.
@@
 Adding and removing the Boltzmann constant from the entropy expression, then, is equivalent to shift from an energy/temperature first approach to an entropy/$\beta$ first approach and should be not seen as physically significant.
+It is really just fixing a unit.
 
@@
 And a macrostate is a probability distribution over microstates.
-It is important to keep these two distributions straight, especially because a microstate can be represented in two ways.
+It is important to distinguish between these two distributions, especially because a microstate can be represented in two ways.
 
@@
 The Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$ is expressed in this space.
-On $\mu$-space, the Shannon entropy $-\int\rho \log \rho$ does not coincide with the thermodynamic entropy for a simple reason: it does not depend linearly with the number of particles.
+On $\mu$-space, the Shannon variability $-\int\rho \log \rho$ does not coincide with the thermodynamic entropy for a simple reason: it does not depend linearly on the number of particles.
 It is not an extensive quantity.
-The reason is straight forward: the Shannon entropy works on normalized distributions and gives us the variability of a single element.
+The reason is straightforward: the Shannon entropy works on normalized distributions and gives us the variability of a single element.
 In this case, the system is composed by $N$ elements, not one.
-One can use the expression $-N\int\rho \log \rho$ as this represents the variability of $N$ elements.\footnote{This is actually an approximation of the multinomial distribution for large N, which would give the precise result.}
+One can use the expression $-N\int\rho \log \rho$ as this represents the variability of $N$ elements.
+
+We want to stress that a distribution on $\mu$-space can only either describe a single microstate where $N$ is very large and therefore the distribution can be assumed to be continuous, or a probability distribution for a $N$ identically and independently prepared particles.
+It one wanted to describe a macrostate, one would have to work with a probability distribution (i.e.
+the weights represents probabilities) over the space of particle distributions (i.e.
+the weights represent fractions of total particles) over $\mu$-space or one would obtain erroneous results.\footnote{This happens, for example, in [Jaynes] where he creates the distribution over $\mu$-space by averaging over all microstates.
+This essentially blends together the probabilities and particle counts losing most of the actual statistical information.
+Naturally, the entropy thus computed is wrong and found to not work.}
 
@@
 Yet, one has to be careful of two things.
-The first is that permutations of particles within $\Gamma$-space lead to different points while should not lead to different microstates.
+The first is that permutations of particles within $\Gamma$-space lead to different points which should not lead to different microstates.
 This leads to the widely known problem of overcounting which needs to be addressed in the standard way.
@@
 
+\section{Comments on Boltzmann's $H$ Theorem}
+
+[This is for you Gabriele].
+In Boltzmann's 1872 paper, he introduced the function that later was called the $H$ function, though it seems he meant to call it [upper case] $\eta$ and it was printed as $E$ (which somehow later became $H$).
+In the paper Boltzmann describes (in $\mu$-space) a gas of monatomic molecules that has already become spatially uniform and also has the same distribution wrt the particles' kinetic energies $x$ at each space point.
+He therefore considers a spatially constant distribution function $f(x,t)$ and wants to show that whatever its initial value it will tend in accordance with the Boltzmann equation (which he has previously derived and I won't give here) to the Maxwell distribution $f(x,t)=C\sqrt xe^{-hx}$ as $t\rightarrow\infty$.
+He defines his function $E$ as follows:
+\begin{equation}
+E=\int_0^\infty f(x,t)\left\{\textrm{log}\left [{f(x,t)\over \sqrt x}\right ]-1\right\}\textrm dx\label{hfun}
+\end{equation}
+and aims to prove that it cannot increase and attains its least value when $f(x,t)$ becomes the Maxwell distribution.
+Boltzmann does not give the argument that leads him to the function (\ref{hfun}), but somewhat later in the paper he notes that the $-1$ in the integrand merely reflects the fact that
+$$
+\int_0^\infty f(x,t)\textrm dx
+$$
+is the total number of molecules in unit volume, so that 
+\begin{equation}
+E=\int_0^\infty f(x,t)\textrm{log}\left [{f(x,t)\over \sqrt x}\right ]\textrm dx\label{hfunb}
+\end{equation}
+also cannot increase.
+
+Boltzmann does not say how he arrived at the function (\ref{hfun}), but its form is obviously dictated by the form of the Maxwell distribution.
+I'm terrible at equations, but I assume the $\sqrt x$ in the denominator of the integrand in (\ref{hfunb}) is because the usual Shannon expression is `skewed' by the subsidiary Lagrangian multiplier that fixes the total energy to a definite value.
+Later in the paper, when considering polyatomic molecules, he arrives by means of arguments I have not yet attempted to understand at the familiar $f\, \textrm{log}\,f$ form.
+
+In his very concise 1860 derivation of his energy distribution function, Maxwell commented that
+
+\begin{quote}\small
+
+It appears from this proposition that the velocities are distributed among the particles according  to the same law as the errors are distributed among the observations in the theory of the ``method of least squares''.
+
+\end{quote}\normalsize
+
+In his much more substantial 1867 paper, Maxwell noted that his 1860 derivation could be seen as `precarious' and gave an alternative proof.
+However, this does not eliminate the mathematical identity of his distribution with the error distribution.
+Would you agree that this (somewhat indirectly) establishes another example of variablity within a distribution of things [which I think we might use instead of `elements'; `members' (of a set) is another word we might consider]?
+
 \section{Conclusion}

### 2019-09-26 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Changing entropy to variability

@@
 It may represent the fraction of elements if the set is constituted by a group of objects.
-It may represent a probability of an outcome is the set is constituted by different takes of similarly prepared systems.
+It may represent a probability of an outcome if the set consists of different realizations of similarly prepared systems.
 If the set is infinite, it will represent some type of convergence.
@@
 
-We note that these are exactly the requirements Shannon put forth for his entropy\cite{Shannon}, and that these requirements lead to only one possible option: $H(p_i) = - \sum p_i \log(p_i)$.
+We note that these are exactly the requirements Shannon put forth for his expression\cite{Shannon}, and that these requirements lead to only one possible option: $H(p_i) = - \sum p_i \log(p_i)$.
 
-The Shannon entropy, then, is a clear and important attribute of a distribution: it quantifies the variability of the elements within the distribution.
+The Shannon variability, then, is a clear and important attribute of a distribution: it quantifies the variability of the elements within the distribution.
 It quantifies the variety of values one finds within the distribution.
@@
 But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope a probability and so on.
-And like a real number represents a quantity, transcending what is physically quantified, the Shannon entropy represents the variability of the elements\footnote{We again stress that out elements are physical and therefore posses some property beyond the bare `definite and separate' of Cantor's definition, which was perfectly suited to study numbers.
+And like a real number represents a quantity, transcending what is physically quantified, the Shannon variability represents the variability of the elements\footnote{We again stress that out elements are physical and therefore posses some property beyond the bare `definite and separate' of Cantor's definition, which was perfectly suited to study numbers.
 Without these extra attributes, which are `washed out' by said definition, we would not be able to create a distribution in the first place.} in a distribution, transcending what the distribution physically signifies.
 
-The Shannon entropy may represent uncertainty in some cases, if the weights are probabilities or credences, but not in the general case.
+The Shannon variability may represent uncertainty in some cases, if the weights are probabilities or credences, but not in the general case.
 If the weights portray the fraction of the elements that have a certain property, like the fraction of galaxies in the Laniakea Supercluster that are barred spirals, there is nothing uncertain about the distribution.
-In this case, the Shannon entropy represents how much variation we have within galaxies in terms of galaxy types.
+In this case, the Shannon variability represents how much variation we have within galaxies in terms of galaxy types.
 
-The Shannon entropy may represent knowledge in some cases, but not in general. Consider the following two cases:
+The Shannon variability may represent knowledge in some cases, but not in general. Consider the following two cases:
 \begin{enumerate}
@@
 \end{enumerate}
-The distribution in both cases is the same, two bins 50\% chance each, and so is the Shannon entropy. Yet, you know more in the second case: you know you have won at least half a million dollars.
+The distribution in both cases is the same, two bins 50\% chance each, and so is the Shannon variability. Yet, you know more in the second case: you know you have won at least half a million dollars.
 
-Unfortunately, entropy in general is often associated with vague characterizations like the the two presented.
+Unfortunately, entropy [CAN'T REALLY CHANGE IT HERE] in general is often associated with vague characterizations like the the two presented.
 It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending by the authors[CITATION NEEDED], which leads to confusion and misunderstanding.
 The characterization of the Shannon variability measure we have given, on the other hand, applies to all cases and leads naturally to the assumptions required to rederive it.
-OUR CHARACTERIZATION IS therefore is more fundamental.
+Our characterization therefore is more fundamental.
 If we look at the galaxies in our universe, what variability do they exhibit in terms of their types?
@@
 
-We have seen that the Shannon entropy measures the variability within a distribution but, in the end, it is a numeric quantity and we cannot expect to understand the concept if we don't understand what the number represents and in what units.
+We have seen that the Shannon variability measures the variability within a distribution but, in the end, it is a numeric quantity and we cannot expect to understand the concept if we don't understand what the number represents and in what units.
 
@@
 
-If we have binary questions, the logarithms will be in base two and the unit for Shannon entropy will be bits.
+If we have binary questions, the logarithms will be in base two and the unit for Shannon variability will be bits.
 It will indicate the average number of yes/no questions we need to identify an element within the dsitribution.
@@
 We can also pick a non-integer base, like the natural base for logarithms, and we will have nats.
-This is why the Shannon entropy is fundamental in information theory, because it quantifies how much information is needed to transfer a value from a distribution.
+This is why the Shannon variability is fundamental in information theory, because it quantifies how much information is needed to transfer a value from a distribution.
 
@@
 \end{itemize}
-The average bit for encoding is $(.48) * 1 + (.27) * 2 + (.1 + .08 + .04 + .03) * 4 = 2.02$ bits while the Shannon entropy is 1.98 bits. That is, the coding is very close to the ideal case.
+The average bit for encoding is $(.48) * 1 + (.27) * 2 + (.1 + .08 + .04 + .03) * 4 = 2.02$ bits while the Shannon variability is 1.98 bits. That is, the coding is very close to the ideal case.
 
@@
 
-The Shannon entropy of the distribution, then, has nothing to do with the information the distribution itself holds.
+The Shannon variability of the distribution, then, has nothing to do with the information the distribution itself holds.
 The distribution is not what is being encoded.
-The entropy is the information needed to go from the distribution, which is given, to an individual element.
+The variability is the information needed to go from the distribution, which is given, to an individual element.
 It's really the information gap from the population to an element.
-That is why some people say the entropy is ``lack of information", because it's the information the distribution does not have.
+That is why some people say the entropy [CAN'T CHANGE IT HERE] is ``lack of information", because it's the information the distribution does not have.
 But, again, this is just missing the point of what is being described.

### 2019-10-07 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Another pass at the language

@@
 It certainly would not admit scientific investigation.
-One of us recalls but has been unable to relocate a sentence in one of Leibniz's many writings to the effect that reality \emph{is} variety.
+We recall, though we have been unable to relocate, a sentence in one of Leibniz's many writings to the effect that reality \emph{is} variety.
 Whether he made the statement or not, it is consistent with his view that any contingent thing is defined by a collection of attributes, which necessarily must be distinguishable
@@
 Indeed, as we will show in the next section, opportunities for application of Shannon's formula are ubiquitous both on the earth and in the universe at large.
-PRECISELY BECAUSE OF THIS GENERALITY WE WILL OFTEN USE THE EXPRESSION SHANNON `FORMULA' RATHER THAN `ENTROPY' OR `INFORMATION'.
 
@@
 This leads us to propose a different characterization of Shannon's formula that is both more general and more appropriate.
-Indeed, we will argue that the formula is a measure of the \emph{variability within a distribution of things of a given kind}.
+Indeed, we will argue that the formula is a measure of the \emph{variability of the values within a distribution of a given set of elements} and we will use the expression Shannon ``variability'', rather than ``entropy'' or ``information''.
 Moreover, the formula has a very wide range of applicability because the universe contains many different kinds of things that, through having common attributes, can be recognized as belonging together despite the fact that their attributes are expressed to different degrees or in different ways.
@@
 We take it as a given.
-
-NEW PARAGRAPH
 
@@
 They were the only attributes that Cantor needed; we need more.
-
-\vspace{.5in}
-
-TODO: Introduction and Conclusions still need to be done
-
-Current disputed status of Shannon entropy as a physical object.
-
-Immense regularity in nature leads to categorization of elements intro distributions.
-Stress that these distributions may be of different nature?
-I.e.: statistical distributions of elements that exist at the same time, probability distributions of events that may happen given similar circumstances, credence distributions of expectation for an agent that something will happen.
-May be better to have a another word instead of "probability" when talking about the elements of the distribution.
-Maybe "weight" or "fraction"?
-
-A natural question is to ask how much variability an element has within a distribution.
-Tempted to just count the number of cases over which the distribution is defined.
-But the shape also affects the variability (e.g.
-a distribution concentrated in few values means less variability of the elements than one equally distributed).
-
-In this paper we will show that $- \sum p_i \log p_i$ is the most natural form to describe the variability of the elements of a distribution.
-
-(NOTE: defining the concept like this is actually broader than Shannon entropy, which assumes source/destination/message.)
 
@@
 Though this is source of additional confusion, for the purpose of defining a measure of variability it does not change things conceptually.
-We will discuss the problem in connection to entropies in thermodynamics and statistical mechanics, since coordinate invariance is paramount in physics.}
+We will discuss the problem in connection to entropies in thermodynamics and statistical mechanics, since coordinate invariance is of paramount importance in physics.}
 
@@
 What we would like to say is that the variability of the overall distribution $H_T = H_C + p_M H_M$ will be the variability over the classes $H_C$ plus the variability of the mammals $H_M$ weighted by the fraction of mammals $p_M$.
+This also makes the quantity linear when combining two independent distributions.
 
@@
 
-The Shannon variability, then, is a clear and important attribute of a distribution: it quantifies the variability of the elements within the distribution.
+The Shannon entropy, which we will call Shannon variability, is a clear and important attribute of a distribution: it quantifies the variability of the elements within the distribution.
 It quantifies the variety of values one finds within the distribution.
@@
 
-Unfortunately, entropy [CAN'T REALLY CHANGE IT HERE] in general is often associated with vague characterizations like the the two presented.
+Unfortunately, entropy in general is often associated with vague characterizations like the the two presented.
 It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending by the authors[CITATION NEEDED], which leads to confusion and misunderstanding.
@@
 If we look at the words in the complete works of Shakespeare, what variability we find in his vocabulary?
-All these questions are interesting, all these questions are profound and the answer to all these questions are expressed through just one thing: the Shannon variability measure.
+All these questions are interesting, all these questions are profound and the answer to all these questions are expressed through just one thing: the Shannon variability.
 
@@
 We can gain insight with the following example.
-Suppose we fix a distribution, say the animals in the Galap\'{a}gos islands grouped by their respective species.
+Suppose we fix a distribution, say the animals in the Galap\'{a}gos islands binned by their respective species.
 Suppose we pick a specific animal from the set and you want to know its species.
@@
 It's really the information gap from the population to an element.
-That is why some people say the entropy [CAN'T CHANGE IT HERE] is ``lack of information", because it's the information the distribution does not have.
+That is why some people say the entropy is ``lack of information", because it's the information the distribution does not have.
 But, again, this is just missing the point of what is being described.
@@
 This is answered by the Shannon variability, which gives us the number of bits required to identify a microstate given the macrostate.
-Since the macrostates fully defines the distribution of possible microstates, this distribution does not depend on the particular way the macrostate came to pass.
+Since the macrostates fully define the distribution of possible microstates, this distribution does not depend on the particular way the macrostate came to pass.
 It is path independent: it is a state variable.
@@
 What we can say, though, is that all quantities that increase over irreversible processes, are maximized at equilibrium and that are additive for indipendent systems are precisely the quantities that are linear functions of entropy.[CITE AXIOMATIC WORK] A linear function of entropy, though, simply means a redefinition of $S_A$ and a change of units for either $dQ$ or $T$.
-Therefore if we can show that the Shannon variability possesses those three properties than it is equivalent to the entropy up to a change in units, which brings us to the second question.
+Therefore if we can show that the Shannon variability possesses those three properties than it is equivalent to the entropy up to a change in units, which would answer our second question.
 
@@
 The Shannon variability, then, has all the right properties to establish a conceptually meaningful connection with the thermodynamic/statistical mechanics entropy.
+It is a state variable that represents the variability of microstates within a macrostate.
+It increases during irreversible processes, reaches the maximum at equilibrium and it is additive for independent systems.
 It is indeed an independent concept but when applied to macrostates yields a state variable with the right conceptual characteristics.
@@
 The first is the role of the Boltzmann constant $k_B$, which is a conversion factor between the Shannon variability.
-$S = - \sum p_i \log(p_i)$ and the Gibbs entropy $S_G = - k_B p_i \sum \log(p_i)$.
+$H = - \sum p_i \log(p_i)$ and the Gibbs entropy $S_G = - k_B p_i \sum \log(p_i)$.
 This conversion constant should not surprise us, since we only established a linear relationship between variability and entropy.
-The question is whether this conversion itself is has a physical meaning, or it is also simply due to a choice of units.
+The question is whether this conversion constant itself has a physical meaning, or it is also simply due to a choice of units.
 The standard relationship between temperature, internal energy and entropy is $T = \frac{\partial U}{\partial S}$.
@@
 The second problem is that we should not think of the particles as literally pointlike: they would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon entropy.
-It is more appropriate, both mathematically and conceptually, to think of indistinguishable particles as identical distributions over $\mu$-space, each contributing the same amount of Shannon entropy.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these single particle distributions is a natural way to incorporate the effects of the uncertainty principle.}
+It is more appropriate, both mathematically and conceptually, to think of indistinguishable particles as identical peaked distributions over $\mu$-space, each contributing the same amount of Shannon entropy.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.}
 
@@
 The Jacobian determiants are unitary and the Shannon entropy is invariant.
-It is not the invariance of the areas over Hamiltonian evolution that is important: it is the invariance over coordinate transformations.
-Invariance over time would be insignificant if we first didn't have a quantity that was observer independent.
+This brings us to another point of confusion: it is often argued that phase space areas are important because they are conserved under Hamiltonian evolution.
+While they are indeed conserved, that is not the reason why they are important.
+It is their invariance under coordinate transformation that makes them physically significant.
+In fact, conservation over time would not even be well defined if the quantity was not invariant and therefore have different values depending on the observer.
 
 This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\rho \log \rho$.
-If we use distributions over position and velocities or over momentum only, the expression we changed for those particular variables in those coordinate systems.
+If we use distributions over position and velocities or over momentum only, the expression would change for those particular variables in those coordinate systems.
+This can be, again, a source of confusion.
 
-These technical details are important to not get confused, but they do not alter the conceptual point.
-The Shannon entropy and the thermodynamic entropy, though describing different things, are conceptually related.
-The variability of the state of the parts within the state of the whole is what the Shannon entropy quantifies if applied to the macrostates of thermodynamics and the microstates of statistical mechanics.
-This variability increases during irreversible evolution, more microstates converge to the same macrostate until equilibrium is reached.
+These technical details are important to use the Shannon variability correctly, but they do not alter the conceptual point.
+The Shannon variability, the thermodynamic entropy and the Gibbs entropy, though describing different things, are conceptually related.
+The variability of the state of the parts within the state of the whole is what the Shannon variability quantifies if applied to the macrostates of thermodynamics and the microstates of statistical mechanics.
+This variability increases during irreversible evolution, more microstates converge to the same macrostate until equilibrium is reached, where the variability is maximized.
 The variability of one system does not affect the variability of an independent one and therefore it simply sums.

### 2019-10-09 Proceedings/2019-Quantum2019/CarcassiQuantum2019Proceedings.tex
Reviewed proceedings with Christine

--- 2019-09-18 Proceedings/2019-Quantum2019/CarcassiQuantum2019Proceedings.tex
@@
 \begin{abstract}
-In a previous work we have shown how the principal difference between classical Hamiltonian mechanics and quantum mechanics is not the physical size of the object being studied, but the degree of accessibility its internal dynamics.
-Namely, in classical mechanics it is assumed to be fully accessible while in classical mechanics is assumed to be totally inaccessible.
-As information entropy can be used to characterize how much the state of the whole system characterizes the state of its parts, it provides fundamental insights that allow us to properly understand the analogies and differences between the two theories.
-Conservation of information, for example, leads to Hamiltonian and Schroedinger evolutions, which describe the deterministic and reversible dynamics.
-We review
+We show that the main difference between classical and quantum systems can be understood in terms of information entropy.
+Classical systems can be considered the ones where the internal dynamics can be known with arbitrary precision while quantum systems can be considered the ones where the internal dynamics cannot be accessed at all.
+As information entropy can be used to characterize how much the state of the whole system identifies the state of its parts, classical systems can have arbitrarily small information entropy while quantum systems cannot.
+This provides insights that allow us to understand the analogies and differences between the two theories.
 \end{abstract}
@@
 By assumptions here we mean simplifying conditions that our system is supposed to satisfy under the processes under study.
-Our work differs from others in that the starting points are physical assumptions about the system being studied (e.g whether the parts can be studied or whether the evolution is deterministic) and not some abstract mathematical or information theoretic considerations.
+Our work differs from others\cite{PhysRevA.84.012311} CITE in that the starting points are physical assumptions about the system being studied (e.g whether the parts can be studied or whether the evolution is deterministic) and not information theoretic or abstract mathematical  considerations.
 It is also different in that both classical and quantum cases are derived on equal footing.
-That is, we have identified a single point where one assumption leads to the classical case while another lead to the quantum system.
+That is, we have identified a single point where one assumption leads to the classical case while another leads to the quantum system.
 
@@
 If it is fully accessible we recover classical states and if it is totally inaccessible we recover quantum states.
-This difference can be understood in terms of information entropy and in this paper we develop that insight.
+This difference can be understood in terms of information entropy and in this work we develop that insight.
 We will present here a much shorter version of the classical derivation that carries all the important points.
-For quantum mechanics, unfortunately we do not currently have one, so we will cover only some of the general ideas.
-For details on both, we refer to the previous longer paper.
+For quantum mechanics, for the sake of brevity, we will cover only some of the general ideas.
+For details on both, we refer to the previous longer paper\cite{Carc1}.
 We then provide a classical analogue of the quantum uncertainty relationship and see how both can be understood in terms of information entropy conservation under the different initial assumptions.
@@
 We define a classical system as one that is assumed to be infinitesimally reducible.
-That is, the system can be divided into parts indefinitely and giving the state of the system is equivalent to giving the state of its parts.
+That is, the system can be thought as made of indefinitely small parts and giving the state of the system is equivalent to giving the state of its parts.
 For example, we can throw a ball and study its motion; or we can take a red marker, make a dot on the ball and study the motion of the red dot.
@@
 over the particle state space $\mathcal{S}$.
-If we have a set of state variable $\xi^a : \mathcal{S} \to \mathbb{R}$ that are sufficient to characterize our particles, we can write both the density and the information entropy
+If we have a set of state variables $\xi^a : \mathcal{S} \to \mathbb{R}$ that are sufficient to characterize our particles, we can write both the density and the information entropy
 \begin{equation}
@@
 
-The issue is that changing coordinates system will change the numeric value of the density.
-Yet, the density is fully specified by the particle states and, since states are invariant under coordinates transformations, it should be invariant.
+The issue is that changing coordinate system will change the numeric value of the density.
+Yet, the density is fully specified by the particle states and, since states are invariant under coordinate transformations, it should be invariant.
 That is, we have
@@
 \end{equation}
-which means the Jacobian of the transformation are unitary.
+which means the Jacobian of the transformation is unitary.
 The invariance of areas is equivalent to the invariance of densities and of information entropy.
@@
 This gives us Hamilton's equations.
-This tells us that the following four concepts
+This tells us that the following four concepts are very tightly related:
 \begin{itemize}
@@
 	\item Conservation of information entropy - the information required to describe the system does not change in time
-	\item Conservation of energy - Hamiltonian evolution
+	\item Conservation of energy - Hamiltonian evolution.
 \end{itemize}
-are the same concept from different angle.
 
@@
 Conversely, we define a quantum system as one that is assumed to be irreducible.
-That is, the state of the system tells us nothing about the state of the parts.
+That is, the system can still be thought as made of parts but the state of the system tells us nothing about the state of the parts.
 For example, we can isolate an electron and study its motion; yet we cannot put a red mark on part of the electron, we cannot scatter a photon off of only a part of an electron.
@@
 It will exhibit diffraction and interference patterns, have half-integer spin and so on.
-If we interact in a way that disturbs the internal dynamic, however, we will see that there is internal dynamics in terms of quarks and gluon.
-In those regimes the proton cannot be treated a single quantum system, but as a composite system.
+If we interact in a way that disturbs the internal dynamics, however, we will see that there is internal dynamics in terms of quarks and gluon.
+In those regimes the proton cannot be treated as a single quantum system, but must be treated as a composite system.
 
-We weren't able to prepare a shortened version of the quantum derivation, but the gist is that, since the state of the system does not capture the state of the parts, we can imagine these continuously permutating without affecting the dynamic of the whole.
-The internal motion is the characterized by random variables, a pair $(A,B)$ for each degree of freedom, which maximize the entropy of the distribution of the parts.
-This means that the variable along each degree of freedom are independent and only the second moments are significant (i.e.
-statistically equivalent to a multivariate Gaussian).
-A complex number $\sigma_A + \imath \sigma_B$, then, can be used to keep track of a variable within a degree of freedom and ultimately leads to the state space of quantum mechanics.
-The norm of the quantum number represents the strength of the internal random process, which is proportional to the size of the system, and the phase is the correlation angle.
-Superposition, then, means superimposing the random motions which will interfere constructively where correlated and interfere destructively where anti-correlated.
+To summarize the arguments laid out in \cite{Carc1}, since the state of the system does not capture the state of the parts, we can imagine these continuously permutating without affecting the dynamics of the whole.
+The internal motion is then characterized by random variables, which still need to allow for the definition of an invariant density, and therefore they must, for a degree of freedom, form a pair $(U,V)$.
+As the value of the variable is completely unknown, the distribution will be uniform.
+We can fix the density $\rho=1$ to be unitary and the standard deviations $\sigma_U=\sigma_V$ to be equal as this internal relabelling cannot affect the state of the system as a whole.
+The transformations that will affect the system as a whole are the changes in variance and correlation as these would change the value of the integral $\int \rho dudv \propto \sigma_U\sigma_V = \sigma_U^2$.
+These are the transformations of the type:
+\begin{equation}
+	\begin{aligned}
+	\hat{U} = a U + b V \\
+	\hat{V} = -b U + a V
+	\end{aligned}
+\end{equation}
+A complex number $c=a + \imath b$ can be used to identify such transformations.
+The norm represents the strength of the internal random process, which is proportional to the size of the system, and the phase relates to the correlation as the arccosine of the Pearson coefficient.
 
@@
 A point-particle is the limit where the extent of the distribution can be considered small, and therefore the center of mass is enough for our description.
-This limit can be performed within both classical and quantum mechanics if the scale of the problem allows, and, because of Ehrenfest theorem, the behavior is the same.
-In light of the assumptions, it should be obvious this should be the case: the difference between classical and quantum is in the description of the parts; if the length scale are much greater than the system, then it doesn't matter what we assume about the parts.
+This limit can be performed within both classical and quantum mechanics if the scale of the problem allows, and, because of Ehrenfest's theorem, the behavior is the same.
+In light of the assumptions, it should be obvious this should be the case: the difference between classical and quantum is in the description of the parts; if the length scales are much greater than those of the system, then it doesn't matter what we assume about the parts.
 
@@
 
-Since classical Hamiltonian evolution conserves information entropy, the space of distribution explored by time evolution is limited by that constrain.
+Since classical Hamiltonian evolution conserves information entropy, the space of distributions explored by time evolution is limited by that constraint.
 We then ask the following question: what is the distribution that, given a certain amount of information entropy, minimizes the spread over phase space?
 That is, what is $\rho(q, k)$ such that $\sigma_q\sigma_k$ is minimized?
-Given the constrain, no Hamiltonian evolution can lead to a state less spread than that.
+Given the constraint, no Hamiltonian evolution can lead to a state less spread than that.
 
@@
 \begin{equation}
-I_G = 2 \pi e \sigma_q \sigma_k.
+I_G = \ln (2 \pi e \sigma_q \sigma_k).
 \end{equation}
@@
 \begin{equation}
-\sigma_q \sigma_k \leq \frac{I_0}{2 \pi e}
+\sigma_q \sigma_k \leq \frac{e^{I_0}}{2 \pi e}
 \end{equation}
 where $I_0$ is the initial information entropy.
-This classical uncertainty relationship is reminiscent of the quantum uncertainty relationship for both the form and the fact that the equality holds for the Gaussian packages.
-The analogy can be made stronger with a few extra conditions, though we will not expand that here.
+This classical uncertainty relationship is reminiscent of the quantum uncertainty relationship in both the form and the fact that the equality holds for the Gaussian packets.
+The analogy can be made stronger with a few extra conditions as described in Ref\cite{Carc1}.
 What we want to show here is the connection between this uncertainty relationship, information entropy, deterministic and reversible evolution and the difference in reducibility between a classical and a quantum system.
@@
 First of all we note that this classical uncertainty relationship is given by conservation of entropy and therefore deterministic and reversible evolution.
-We are free to set the initial amount of information entropy for the classical distribution.
+We are free to set the initial amount of information entropy for the classical distribution as small as we want.
 Once it is set, this cannot be changed under Hamiltonian evolution.
-Quantum mechanics has a stronger condition: it is not possible to create a state that violates the uncertainty.
-That is, we do not have the initial freedom we have in classical mechanics.
+Quantum mechanics has a stronger condition: we cannot set the initial uncertainty as small as we want.
+It is bound by a fundamental constant.
 For quantum states, all pure states have the same information entropy which is set to zero.
@@
 
-The deterministic and reversible case in quantum mechanics is given by unitary evolution, which in fact conserves Von Neumann entropy because each pure state is mapped to a pure state.
-Conversely, the Lindblad equation allows the entropy to increase and corresponds to irreversible evolution.
-However, in that case it is the entropy of the mixed case that is being increased.
-In classical mechanics this would correspond to mapping one initial distribution to multiple final distribution with the same entropy according to different probability.
-That is, we have a probability distribution over matter distributions.
-As each matter distribution has the same entropy, it is the spread in probability that increases the entropy.
-But in classical mechanics we can also do something more: we can have a density distribution spread more or less over phase space with a dynamic that is not Hamiltonian and change the information entropy of the matter distribution itself.
-For example, a damped harmonic oscillator would concentrate the distribution around the equilibrium and decrease the entropy.
-This type of dynamic cannnot be described in the context of quantum mechanics.
-Another aspect of this same issue is that given two quantum states we can always find a unitary evolution that connects the two (at least mathematically) while this is not possible in classical Hamiltonian mechanics (e.g.
+The deterministic and reversible case in quantum mechanics is given by unitary evolution, which in fact conserves von Neumann entropy because each pure state is mapped to a pure state.
+Moreover, given two pure quantum states of the same system we can always find a unitary evolution that connects the two (at least mathematically) while this is not possible in classical Hamiltonian mechanics (e.g.
 two distributions with different entropy cannot be linked by Hamiltonian evolution).
 
-We believe these insights clarify two issues.
-The first is that the difference between a classical and quantum system is not primarily about is size.
-The previous discussion does not depend on the physical extension of the system or about the number of its degrees of freedom.
+These insights lead us to conclude that the difference between a classical and quantum system is not primarily about its size.
+The previous discussion does not depend on the physical extent of the system or the number of its degrees of freedom.
 It is about the accessibility of the internal dynamics.
 If we are to look at scenarios where classical mechanics fails, we simply have to look at cases where we can no longer assume that the internal dynamics is indefinitely accessible.
-In a Bose-Einstein condensate, in a pair of entangled photons or in a diffracting neutron the parts cannot be studied independently without significantly altering the system: infinitesimal reducibility does not apply.
-Conversely, if we are to look at scenarios where quantum mechanics fails as applied for a specific system, we simply have to look at cases where the internal dynamic of that system is indeed accessible.
-
-The second issue is that it tells us what is being quantized to begin with.
-Quantities like energy, spin or momentum are indeed quantized in some cases, but they are not always quantized in the same way and there is a continuum of different cases.
-For example, spin along one direction is quantized, but we can continuously change the direction.
-Energy is quantized only for bound systems, and even in those cases a relative boost can continuously change the value.
-The entropy associated with all pure states, instead, is always the same for all states in all frames.
-Again, this is not true in classical mechanics: a matter distribution over phase space can have an arbitrary information entropy.
-The entropy associated to a pure quantum state, which is itself a distribution (i.e.
-a wave function), is always the same.
-That is what is quantized.
-It if weren't, the Von Neuman expression would not be the correct expression for entropy as each pure state would have to be weighted different.
-The lack of that term means all pure states carry the same contribution: the information entropy of pure states is what is quantized.
-
-\section{Submitted abstract}
-In a recent paper [1] we have shown that classical and quantum Hamiltonian particle mechanics can  be  rederived  from  a  handful  of  physical  assumptions.
-We  will  present  and  expand  a consequence   of   that   work,   namely   that   classical   Hamiltonian   mechanics   coincides   with conservation of information entropy, and how this insight allows us to better understand quantum mechanics as well.We  will  start  by  showing  that  classical  Hamiltonian  mechanics  can  be  recovered  from  two fundamental assumptions.
-The first is that the system we study is infinitesimally reducible: it is made  of  infinitesimal  parts  and  the  state  of  those  parts  is  equivalent  to  the  state  of  the  whole system.
-The state of the overall system is then a distribution over the states of the infinitesimal parts.
-As states must be defined independently of coordinate system (i.e.
-of observer), the density over the states of the parts must be invariant under coordinate transformation.
-This allows us to recover  the  geometrical  structure  of  phase  space  (i.e.
-state  variables  organized  in  conjugate pairs).
-The second assumption is that of deterministic and reversible evolution: knowing the state at one time is enough to predict the state at future times and reconstruct the state at past times.
-This means that the density for a given state will be mapped exactly to a future state, which will give us Hamiltonian evolution.
-We then show that requiring densities to be coordinate invariant and  not  change  in  time  is  equivalent  to  requiring  that  the  information  entropy  (i.e.
-the  number bits of identify an element of the distribution up to unit precision) is conserved.Throughout the derivation, we will see how classical analogues of quantum mechanics (i.e.
-wave number, uncertainty principle, particle/anti-particle states) naturally emerge.
-We will discuss how quantum mechanics differs from classical mechanics by assuming irreducibility instead: that the system is still made of infinitesimal parts but that the state of the overall system tells us nothing about  the  state  of  the  parts.
-We  will  see  how  this  qualitatively  leads  to  the  known  quantum effects.This work is part of a larger project [2] that aims to rederive the known laws of physics from a handful of physical assumptions.
+Conversely, if we were to look at scenarios where quantum mechanics fails as applied to a specific system, we would need to look at cases where the internal dynamics of that system becomes at least minimally accessible.
 
@@
 
-We have shown that the principal difference between classical and quantum systems is how much the state of the whole can say about it parts.
-This can be understood in terms of information entropy: classical states as distribution over phase space allow for ever decreasing information entropy while quantum states has wave functions have a fixed entropy which is set to zero.
-Deterministic and reversible evolution conserve information, therefore we have an uncertainty relationship under Hamiltonian evolution.
-
-We conclude by noting that both the classical assumptions (i.e.
-the internal dynamic is infinitesimally accessible) and the quantum assumption (i.e.
-the internal dynamic is totally inaccessible) seem untenable.
-If we were to suppose that the internal dynamics, though not influenceable, had consequences (e.g.
-they would decide when a particle decays or how an entangled system is broken), then we would also have processes that increased entropy over time (i.e.
-more information is needed to identify future state).
-Such a change of assumption, though, is not trivial as it would necessarily require rethinking the state space as well.
+We have shown that the principal difference between classical and quantum systems is how much the state of the whole can characterize its parts.
+In classical systems the internal dynamics is fully accessible while in a quantum system it is completely inaccessible.
+This can be understood in terms of information entropy: classical states as distributions over phase space allow for arbitrarily small amounts of information entropy while pure quantum states have a fixed entropy which is set to zero.
 
@@
 
-This section should come before the Appendices and References. Funding
-information may also be included here.
+ Funding for this work was provided in part by the MCubed program of theUniversity of Michigan.
 

### 2019-10-16 Proceedings/2019-Quantum2019/CarcassiQuantum2019Proceedings.tex
Adding bibliography

@@
 By assumptions here we mean simplifying conditions that our system is supposed to satisfy under the processes under study.
-Our work differs from others\cite{PhysRevA.84.012311} CITE in that the starting points are physical assumptions about the system being studied (e.g whether the parts can be studied or whether the evolution is deterministic) and not information theoretic or abstract mathematical  considerations.
+Our work differs from others, see for example Refs.~\citenum{PhysRevA.84.012311,QLogicReview,Hardy:2001jk,ludwig_hein_2013}, in that the starting points are physical assumptions about the system being studied (e.g.
+whether the parts can be studied or whether the evolution is deterministic) and not information theoretic or abstract mathematical considerations.
 It is also different in that both classical and quantum cases are derived on equal footing.
@@
 This classical uncertainty relationship is reminiscent of the quantum uncertainty relationship in both the form and the fact that the equality holds for the Gaussian packets.
-The analogy can be made stronger with a few extra conditions as described in Ref\cite{Carc1}.
+The analogy can be made stronger with a few extra conditions as described in Ref.~\citenum{Carc1}.
 What we want to show here is the connection between this uncertainty relationship, information entropy, deterministic and reversible evolution and the difference in reducibility between a classical and a quantum system.

### 2019-10-22 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Adding Boltzmann entropy and bibliography

@@
 If we fix a macrostate and then measure the state of all its microscopic constituents, we will not get the same answer every time.
-This true in the case of non equilibrium (e.g.
+This is true in the case of non-equilibrium (e.g.
 we suddenly compress a gas pushing a piston) and of equilibrium (e.g.
@@
 At this point, it is not apparent why these two quantities defined independently would be related.
-What we can say, though, is that all quantities that increase over irreversible processes, are maximized at equilibrium and that are additive for indipendent systems are precisely the quantities that are linear functions of entropy.[CITE AXIOMATIC WORK] A linear function of entropy, though, simply means a redefinition of $S_A$ and a change of units for either $dQ$ or $T$.
+What we can say, though, is that all quantities that increase over irreversible processes, are maximized at equilibrium and that are additive for indipendent systems are precisely the quantities that are linear functions of entropy (see \cite{Giles}).
+A linear function of entropy, though, simply means a redefinition of $S_A$ and a change of units for either $dQ$ or $T$.
 Therefore if we can show that the Shannon variability possesses those three properties than it is equivalent to the entropy up to a change in units, which would answer our second question.
@@
 Suppose we have an irreversible process, which means we can start from many out-of-equilibrium states which end into a single equilibrium state.
-During the whole process, some of the macro variables are going to stay constant as they characterize the constraint the system is subjected during the evolution.
-For example, if the walls are fixed, the gas will have a set volume; if the walls are insulated, then the average energy will be conserved.\footnote{It can also happen that the irreversible process does have a single possible equilibrium, as in the case of spontaneous symmetry breaking.
-New variables will identify the equilibrium which we can think of new constraints appeared once the initially symmetry is broken.} As different out-of-equilibrium macrostates merge during the evolution, they merge the possible initial conditions that lead to the same macrostate.
+During the whole process, some of the macro variables are going to stay constant as they characterize the constraint the system is subjected during the evolution and will also identify the final equilibrium.
+For example, if the walls are fixed and insulated, the gas will have a set volume and average energy, which will find a unique equilibrium with a given temperature and pressure.\footnote{It can also happen that the irreversible process does not have a single possible equilibrium, as in the case of spontaneous symmetry breaking.
+New variables will identify the equilibrium which we can think of new constraints, appeared once the initially symmetry is broken, thus reducing to the previous case.} As different out-of-equilibrium macrostates merge during the evolution, they merge the possible initial conditions that lead to the same macrostate.
 The variability of microstates within the initial out-of-equilibrium macrostate, then, must be smaller than the variability of the microstates within the final equilibrium state.
-The amount of information needed to identify a microstate at equilibrium allows us to identify the initial microstate across multiple out-of-equilibrium states and therefore is greater than the information needed to identity a microstate in an out-of-equilibrium macrostate.
+The number of bits needed to identify a microstate at equilibrium allows us to identify the initial microstate across multiple out-of-equilibrium states and therefore is greater than the number of bits needed to identity a microstate in an out-of-equilibrium macrostate.
 Irreversible processes increase the Shannon variability within the macrostate, which corresponds to a greater uniformity of the macrostate itself since uniform distributions are the ones that maximize the variability.
@@
 That is, the distribution over microstates are statistically independent.
-This means that the statistical distribution is the product of the join distributions and, in this case, the Shannon entropy is additive.
+This means that the joint distribution is the product of the marginal distributions and, in this case, the Shannon entropy is additive.
 
@@
 
-The first is the role of the Boltzmann constant $k_B$, which is a conversion factor between the Shannon variability.
-$H = - \sum p_i \log(p_i)$ and the Gibbs entropy $S_G = - k_B p_i \sum \log(p_i)$.
+The first is the role of the Boltzmann constant $k_B$, which is a conversion factor between the Shannon variability $H = - \sum p_i \log(p_i)$ and the Gibbs entropy $S_G = - k_B p_i \sum \log(p_i)$.
 This conversion constant should not surprise us, since we only established a linear relationship between variability and entropy.
@@
 Third, when setting up entropy maximization, $\beta$ is quite naturally the Lagrange multiplier associated with energy conservation, therefore mathematically it has a much clearer role.
-Lastly, one can proceed in both thermodynamic and statistical mechanics with just $\beta$, and convert to temperature only when we need a value on the Kelvin scale.
+Lastly, one can proceed in both thermodynamic and statistical mechanics with just $\beta$, and convert to temperature only when a value on the Kelvin scale is needed.
 For these reasons, it seems to us $\beta$ plays the more fundamental role and $k_B$ is the conversion coefficient to go from $\beta$ to temperature measured in Kelvin.
@@
 
-We want to stress that a distribution on $\mu$-space can only either describe a single microstate where $N$ is very large and therefore the distribution can be assumed to be continuous, or a probability distribution for a $N$ identically and independently prepared particles.
-It one wanted to describe a macrostate, one would have to work with a probability distribution (i.e.
+We want to stress that a distribution on $\mu$-space can only either describe a single microstate where $N$ is very large and therefore the distribution can be assumed to be continuous, or a probability distribution for $N$ identically and independently prepared particles.
+A macrostate cannot be described by a single distribution on $\mu$-space as this cannot capture correlations.
+One would need to work with a probability distribution (i.e.
 the weights represents probabilities) over the space of particle distributions (i.e.
-the weights represent fractions of total particles) over $\mu$-space or one would obtain erroneous results.\footnote{This happens, for example, in [Jaynes] where he creates the distribution over $\mu$-space by averaging over all microstates.
+the weights represent fractions of total particles) over $\mu$-space or one would obtain erroneous results.\footnote{Unfortunately this happens, for example, in [Jaynes] where he creates the distribution over $\mu$-space by averaging over all microstates.
 This essentially blends together the probabilities and particle counts losing most of the actual statistical information.
@@
 If we use distributions over position and velocities or over momentum only, the expression would change for those particular variables in those coordinate systems.
-This can be, again, a source of confusion.
+This can be, again, a source of confusion.\cite{Dunkel}
 
-These technical details are important to use the Shannon variability correctly, but they do not alter the conceptual point.
-The Shannon variability, the thermodynamic entropy and the Gibbs entropy, though describing different things, are conceptually related.
+The last issue is the relationship with the Boltzman entropy $S_B= k_B \log W$.
+This is actually captured by another standard way to derive the Shannon variability.
+Suppose we have a sequence $N$ elements, say twenty balls.
+There are $I$ possible types of elements, say three possible colors: red, green and blue.
+Suppose $N_i$ are the elements of each possible types, which means $\sum_{i \in I} N_i = N$.
+Then, given a set of $N_i$, we have $W(N_i) = \frac{N!}{\prod_{i \in I} N_i}$ possible ways to realize that sequence which corresponds to all possible permutations.
+If all permutations are equally likely, then $\log W$ represents the number of bits needed to identify one of the sequences.
+Since there are $N$ elements in the sequence, $\frac{1}{N} \log W$ represents the average number of bits needed to identify one element.
+We now take the limit for large $N$, that is we ask what is the average number of bits needed to identify one element from an infinite sequence given by the weights $p_i = \frac{N_i}{N}$.
+We find $\frac{1}{N} \log \frac{N!}{\prod_{i \in I} N_i} \approx - \sum p_i \log p_i$.
+In other words, the Boltzman entropy reduces to the Shannon entropy when counting permutations within large distributions.
+
+These technical details are important to understand the fine details of the links between the different entropies, they do not alter, and should not distract us, from the main conceptual point.
+The Shannon variability, the thermodynamic entropy, the Boltzmann entropy and the Gibbs entropy, though describing different things, are conceptually related.
 The variability of the state of the parts within the state of the whole is what the Shannon variability quantifies if applied to the macrostates of thermodynamics and the microstates of statistical mechanics.
@@
 \bibliographystyle{unsrt}  
-\bibliography{references}
+\bibliography{bibliography}
 

### 2019-10-25 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Adding Christine's comments

@@
 In accordance with the comments made in our introduction, they can for example represent the galaxies in the Laniakea Supercluster, the animals in the Gal\'{a}pagos islands, the molecules in a box of gas given a particular macrostate, the different outcomes of spin measurement for a given quantum state, or the words in the complete works of Shakespeare.
-The choice of the particular set is driven by the interest and practices of scientists and their fields of study and therefore it is, in this sense, discretional.
+The choice of the particular set is driven by the interests and practices of scientists and their fields of study and therefore it is, in this sense, discretionary.
 Yet, it must be objective in at least one respect: once a choice is made the actual elements are the same for everybody.
@@
 In turn, this will make some choices arguably more appropriate or interesting than others.
-For example, in taxonomy the set of all animals that have feathers may be more interesting than the set of all animals that can fly: the first includes only birds while the second includes most birds, a few mammals and a lot of insects.
+For example, in taxonomy the set of all animals that have feathers may be more appropriate than the set of all animals that can fly: the first includes only birds while the second includes most birds, a few mammals and a lot of insects.
 
 Once a set of elements is chosen, we select a property or a set of properties we want to use to characterize the elements.
-That is, we have a set of possible values $Q=\{q_i\}_{i=1}^I$ and a map $q : E \to Q$ that associates a value for each element.
+That is, we have a set of possible values $Q=\{q_i\}_{i=1}^I$ and a map $q : E \to Q$ that associates a value to each element.
 These can represent the galaxy types, the genus of the animal, the possible states of the molecules, the possible values for spin or the spelling of words.
 This will give us a sequence $\{q(e_\alpha)\}_{\alpha=1}^N$ of the descriptions that are associated with each element in the set.
-Again, the choice of properties is ultimately discretional, grounded on what particular aspects of the elements we are studying in a given case.
+Again, the choice of properties is ultimately discretionary, grounded on what particular aspects of the elements we are studying in a given case.
 Yet, like before, once a choice is made the value for each element is objectively defined.
@@
 It may represent a probability of an outcome if the set consists of different realizations of similarly prepared systems.
-If the set is infinite, it will represent some type of convergence.
-We will therefore call $p_i$ weights, to be clear that we make no commitment on whether we have fractions, frequentist probabilities, Bayesian probabilities or other notions.
+We will therefore call $p_i$ weights, to be clear that we make no commitment as to whether we have fractions, frequentist probabilities, Bayesian probabilities or other notions.
 
-Once we have constructed our distribution, we will want to construct an indicator $S$ that quantifies how much variability the elements exhibit within the distributions.
+Once we have constructed our distribution, we will want to construct an indicator $H$ that quantifies how much variability the elements exhibit within the distribution.
 That is, we want to quantify the degree of diversity that the values can have within the distribution.
-To that end, we want to define some suitable requirements for $S$.
+To that end, we want to define some suitable requirements for $H$.
 
@@
 First, if the values associated with the bins are non-numerical (e.g.
-types of galaxies, words), statistical quantities are not well defined.
+types of galaxies, words), such statistical quantities are not well defined.
 Second, relabeling the values (e.g.
@@
 This tells us that our indicator cannot depend on $Q$ itself, but only on the weights $p_i$.
-That is, we require $H=H(p_i)$.\footnote{Note that for continuous quantities the weights are densities and are affected by the choice of $Q$: the unit is required to specify the numeric value (e.g.
-$1 \% / m$) and this will change under unit transformation (e.g.
-$1000 \% / Km$).
+That is, we require $H=H(p_i)$.\footnote{Note that for continuous quantities the weights are densities and are affected by the choice of $Q$: the unit is required to specify the numeric value (e.g.~$1 \% / $mm) and this will change under unit transformation (e.g.~$1000 \% /$m).
 Therefore, as the weights $p_i$ themselves depend on the unit, the indicator $H(p_i)$ will in general depend on the choice of $Q$.
-Though this is source of additional confusion, for the purpose of defining a measure of variability it does not change things conceptually.
+Though this is a source of additional confusion, for the purpose of defining a measure of variability it does not change things conceptually.
 We will discuss the problem in connection to entropies in thermodynamics and statistical mechanics, since coordinate invariance is of paramount importance in physics.}
@@
 \item Let $p_i$ and $q_j$ be the weights for two distributions respectively over $I$ and $J$ bins.
-Let $r_k$ be the distribution over $K=I+J-1$ bins constructed by expanding the $a$-th bin of the first distribution using the second distribution.
+Let $r_k$ be the distribution over $K=I+J-1$ bins constructed by expanding the $a^{th}$ bin of the first distribution using the second distribution.
 More specifically, let $1 \leq a \leq I$, then $r_k = \{p_1, p_2, ..., p_{a-1}, p_{a}q_1, p_{a}q_2, ..., p_{a}q_J, p_{a+1}, ..., p_I \}$.
@@
 
-The Shannon entropy, which we will call Shannon variability, is a clear and important attribute of a distribution: it quantifies the variability of the elements within the distribution.
+The Shannon entropy, which we will call Shannon variability, is a clear and important attribute of a given distribution: it quantifies the variability of the elements within the distribution.
 It quantifies the variety of values one finds within the distribution.
-It is true that the choice of the elements and binning are not fixed, it is true that the meaning of the weights depends on what the distribution is describing, and therefore that the precise meaning of this variability is context dependent.
-But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope a probability and so on.
-And like a real number represents a quantity, transcending what is physically quantified, the Shannon variability represents the variability of the elements\footnote{We again stress that out elements are physical and therefore posses some property beyond the bare `definite and separate' of Cantor's definition, which was perfectly suited to study numbers.
+It is true that the choices of the elements and binning are not fixed, it is true that the meaning of the weights depends on what the distribution is describing, and therefore that the precise meaning of this variability is context dependent.
+But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope, a probability and so on.
+And like a real number represents a quantity, transcending what is physically quantified, the Shannon variability represents the variability of the elements\footnote{We again stress that out elements are physical and therefore possess some property beyond the bare `definite and separate' of Cantor's definition, which was perfectly suited to study numbers.
 Without these extra attributes, which are `washed out' by said definition, we would not be able to create a distribution in the first place.} in a distribution, transcending what the distribution physically signifies.
@@
 Unfortunately, entropy in general is often associated with vague characterizations like the the two presented.
-It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending by the authors[CITATION NEEDED], which leads to confusion and misunderstanding.
+It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending on the authors[CITATION NEEDED], which leads to confusion and misunderstanding.
 The characterization of the Shannon variability measure we have given, on the other hand, applies to all cases and leads naturally to the assumptions required to rederive it.
@@
 If we look at the molecules in a given macrostate, what variability do they express in terms of their microstates?
-If we look at the words in the complete works of Shakespeare, what variability we find in his vocabulary?
-All these questions are interesting, all these questions are profound and the answer to all these questions are expressed through just one thing: the Shannon variability.
+If we look at the words in the complete works of Shakespeare, what variability do we find in his vocabulary?
+All these questions are interesting, all these questions are insightful and the answers to all these questions are expressed through just one thing: the Shannon variability.
 
-If the Shannon expression is a measure of variability, why is it connected to information? What is it measuring and in what units?
+If the Shannon expression is a measure of variability, how is it connected to information? What is it measuring and in what units?
 
@@
 
-We have seen that the Shannon variability measures the variability within a distribution but, in the end, it is a numeric quantity and we cannot expect to understand the concept if we don't understand what the number represents and in what units.
+We have seen that the Shannon variability measures the variability within a distribution but, in the end, it is a numeric quantity and we cannot expect to understand the concept if we do not understand what the number represents and in what units.
 
@@
 Defining a set of questions means choosing an encoding as we are choosing how the information gets codified into our series of bits.
-To understand how this works, we can briefly review the Hauffman coding, which is the optimal algorithm for symbol-by-symbol coding with a known probability distribution.
+To understand how this works, we can briefly review the Huffman coding,[CITE] which is the optimal algorithm for symbol-by-symbol coding with a known probability distribution.
 The idea is that we want all possible answers to each question to be balanced, to provide the same amount of information.
-The reason is that making one answer more specific (i.e.
-it applies in fewer cases) means making another less specific (i.e.
-it applies in more cases).
-In the case of binary questions, then, we want the probability to answer yes or no to be 50\%.
+The reason is that making one answer more specific (i.e.~it applies in fewer cases) means making another less specific (i.e.~it applies in more cases).
+In the case of binary questions, then, we ideally want the probability to answer yes or no to be 50\%.
 
-For example, suppose we have a population of pets in a country is as follows:
+For example, suppose the population of pets in a country is as follows:
 \begin{itemize}
@@
 We can group fish with reptile and bird with small mammal to form a 13/14 split.
-So we can ask ``is it a fish or a reptile?" If the answer is yes, the followup question would be ``is it a fish?" If the answer is no, the followup question would be ``is it a bird?" With this scheme, the enconding is as follows:
+So we can ask ``is it a fish or a reptile?" If the answer is yes, the followup question would be ``is it a fish?" If the answer is no, the followup question would be ``is it a bird?" With this scheme, the encoding, where 1 represents 'yes' and 0 represents 'no' to the each of the questions asked, is as follows:
 \begin{itemize}
-\item dogs 27\% - 01 - 2 bits
-\item cats 48\% - 1 - 1 bit
-\item fish 10\% - 0011 - 4 bits
-\item bird 8\% - 0001 - 4 bits
-\item small mammal 4\% - 0000 - 4 bits
-\item reptile 3\% - 0010 - 4 bits
+\item dogs 27\% - 2 questions (i.e.~2 bits) - answers: [no, yes] (i.e.~encoding 01)
+\item cats 48\% - 1 question (i.e.~1 bit) - answers: [yes] (i.e.~encoding 1)
+\item fish 10\% - 4 questions (i.e.~4 bits) - answers: [no, no, yes, yes] (i.e.~encoding 0011)
+\item bird 8\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, yes] (i.e.~encoding 0001)
+\item small mammal 4\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, no] (i.e.~encoding 0000)
+\item reptile 3\% - 4 questions (i.e.~4 bits) - answers: [no, no, yes, no] (i.e.~encoding 0010)
 \end{itemize}
-The average bit for encoding is $(.48) * 1 + (.27) * 2 + (.1 + .08 + .04 + .03) * 4 = 2.02$ bits while the Shannon variability is 1.98 bits. That is, the coding is very close to the ideal case.
+The number of questions needed in each case corresponds to the number of bits.
+The answers in each case are represented by the encoding.
+Note how the encoding depends on the specific choice of questions.
+We can calculate the average number of bits for the encoding to be:  $(.48) * 1 + (.27) * 2 + (.1 + .08 + .04 + .03) * 4 = 2.02$ bits.
+This represents the average number of bits we would have to use for each animal if we repeated the game many times.
+We can also calculate the Shannon variability to be $.27 * \log_2(.27) + .48 * \log_2 (.48) + ...
+=1.98$ bits.
+This represents the ideal case, the minimum number of questions required to reach a definite answer.
+Note that our encoding is already very close to the ideal case.
 
@@
 The variability is the information needed to go from the distribution, which is given, to an individual element.
-It's really the information gap from the population to an element.
-That is why some people say the entropy is ``lack of information", because it's the information the distribution does not have.
+It is really the information gap from the population to an element.
+That is why some people say the entropy is ``lack of information", which is justified because, in a way, it is the information about the elements that the distribution cannot provide.
 But, again, this is just missing the point of what is being described.
@@
 In communication and information theory, information is really encoded information.
-Communication systems and information processors have no idea whether the source of the data is a digital thermometer or a pressure sensor.
+Communication systems and information processors have no idea whether the source of the data is a digital thermometer or a poet.
 There is no knowledge per se, just symbol manipulation that may represent different concepts in different contexts.
 
-The term information entropy, then, is unfortunate for two reasons.
-First, because is not really entropy in the thermodynamic sense: it is not defined on states, it does know about irreversible processes and it is unrelated to maximization at equilibrium.
+The term information entropy, then, is misleading for two reasons.
+First, because it is really not entropy in the thermodynamic sense: it is not defined on states, it does not know about irreversible processes and it is unrelated to maximization at equilibrium.
 In fact, it is not related to physical systems.
-Second, it is not really information in the general sense, only in the very narrow technical sense of coded information within a communication or information system.
+Second, it is not really information in the general sense, only in the very narrow technical sense of encoded information within a communication or information system.
 
-So why is it so successfully used in statistical mechanics? Why does it seem to play a fundamental role even for systems far from equilibrium?
+So why is it so successfully used in statistical mechanics? Why does it seem to play a fundamental role even for systems far from equilibrium?[CITE]
 
@@
 As we saw in the previous sections, Shannon variability is a very general concept that applies to any type of distribution, no matter what the problem domain.
-It does not require to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamic and statistical mechanics which requires them.\footnote{For this reason it is unfortunate that it was given the same name.} It is only when variability is appropriately applied to this particular setting that, though not apparent at first, a connection can be made.
+It does not require us to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamic and statistical mechanics which requires them.\footnote{For this reason it is unfortunate that it was given the same name.} It is only when variability is appropriately applied to this particular setting that, though not apparent at first, a connection can be made.
 To understand this link we need to answer two basic questions.
 First, what does the Shannon variability represent in the context of statistical mechanics and thermodynamics?
-In particular, why is it a state variable and what property of the system is it describing?
+In particular, why can it be treated as a state variable and what property of the system is it describing?
 Second, why does the Shannon variability satisfy the same properties as the entropy?
@@
 If we fix a macrostate and then measure the state of all its microscopic constituents, we will not get the same answer every time.
-This is true in the case of non-equilibrium (e.g.
-we suddenly compress a gas pushing a piston) and of equilibrium (e.g.
-we give enough time to equilibrate).
+This is true in the case of non-equilibrium (e.g.~we suddenly compress a gas by pushing a piston) and of equilibrium (e.g.~we give enough time to equilibrate).
 The choice of a macrostate, then, identifies a set of possible microstates that are compatible with it.
@@
 It is path independent: it is a state variable.
-In short, as a macrostate is a distribution over microstates, it will admit as a state variable the Shannon variability, regardless of whether we know the actual distribution.
+In short, as a macrostate is a distribution over microstates, it will admit as a state variable the Shannon variability, regardless of whether we know the actual distribution and can calculate its value.
 
-Note, though, that the thermodyanmic entropy is defined differently.
+Note, though, that the thermodynamic entropy is defined differently.
 It is defined over a succession of equilibrium states by noting that the integral of the heat exchanged divided by the temperature is path independent.
-One then chooses a particular state $A$, assigns it an arbitrary value $S_A$ and the entropy for all other states is given by the integral $S(B) = S_A + \int_A^B \frac{dQ}{T}$.
+One then chooses a particular state $A$, assigns it an arbitrary entropy value $S_A$ and the entropy for all other states is given by the integral $S(B) = S_A + \int_A^B \frac{dQ}{T}$.
 At this point, it is not apparent why these two quantities defined independently would be related.
-What we can say, though, is that all quantities that increase over irreversible processes, are maximized at equilibrium and that are additive for indipendent systems are precisely the quantities that are linear functions of entropy (see \cite{Giles}).
+What we can say, though, is that all quantities that increase over irreversible processes, are maximized at equilibrium and that are additive for independent systems are precisely the quantities that are linear functions of entropy (see \cite{Giles}).
 A linear function of entropy, though, simply means a redefinition of $S_A$ and a change of units for either $dQ$ or $T$.
@@
 
-Suppose we have an irreversible process, which means we can start from many out-of-equilibrium states which end into a single equilibrium state.
-During the whole process, some of the macro variables are going to stay constant as they characterize the constraint the system is subjected during the evolution and will also identify the final equilibrium.
-For example, if the walls are fixed and insulated, the gas will have a set volume and average energy, which will find a unique equilibrium with a given temperature and pressure.\footnote{It can also happen that the irreversible process does not have a single possible equilibrium, as in the case of spontaneous symmetry breaking.
-New variables will identify the equilibrium which we can think of new constraints, appeared once the initially symmetry is broken, thus reducing to the previous case.} As different out-of-equilibrium macrostates merge during the evolution, they merge the possible initial conditions that lead to the same macrostate.
+Suppose we have an irreversible process, which means we can start from many out-of-equilibrium states which end in a single equilibrium state.
+During the whole process, some of the macrostate variables are going to stay constant as they characterize the constraints to which the system is subjected during the evolution and will also identify the final equilibrium.
+For example, if the walls are fixed and insulated, the gas will have a set volume and average energy, which will lead to a unique equilibrium with a given temperature and pressure.\footnote{It can also happen that the irreversible process does not have a single possible equilibrium, as in the case of spontaneous symmetry breaking.
+New variables will identify the equilibrium which we can think of as new constraints appearing once the initial symmetry is broken, thus reducing to the previous case.} As different out-of-equilibrium macrostates merge during the evolution, they merge the possible microstates that lead to the same macrostate.
 The variability of microstates within the initial out-of-equilibrium macrostate, then, must be smaller than the variability of the microstates within the final equilibrium state.
-The number of bits needed to identify a microstate at equilibrium allows us to identify the initial microstate across multiple out-of-equilibrium states and therefore is greater than the number of bits needed to identity a microstate in an out-of-equilibrium macrostate.
 Irreversible processes increase the Shannon variability within the macrostate, which corresponds to a greater uniformity of the macrostate itself since uniform distributions are the ones that maximize the variability.
 
-To understand why this is maximized at equilibrium, suppose it weren't.
-That is, suppose there was another macrostate which satisfied the same constraints, but for which the Shannon variability was greater.
+To understand why this is maximized at equilibrium, suppose it were not.
+That is, suppose there were another macrostate which satisfied the same constraints, but for which the Shannon variability were greater.
 If we prepared such a state and let it equilibrate, it could not evolve to the equilibrium because that has lower Shannon variability.
@@
 We find this approach, sometimes called entropy first, preferable for a number of reasons, especially in statistical mechanics.
-First, entropy is invariant under coordinate transformations and therefore it is better to keep it a pure number and not mix it with units of energy, which is not invariant under boost.\footnote{This is essential if one ever hopes to put statistical mechanics in a relativistic setting.} Second, the maximization of entropy at equilibrium guaranties us that $S(U)$ is a single valued function (i.e.
+First, entropy is invariant under coordinate transformations and therefore it is better to keep it a pure number and not mix it with units of energy, which is not invariant under boost.\footnote{This is essential if one ever hopes to put statistical mechanics into a relativistic setting.} Second, the maximization of entropy at equilibrium guaranties us that $S(U)$ is a single valued function (i.e.
 the maximum given the value of $U$) and $\frac{\partial S}{\partial U}$ is therefore well defined.

### 2019-10-25 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Updated the introduction

@@
 
-A world without any variety is barely conceivable.
-It certainly would not admit scientific investigation.
-We recall, though we have been unable to relocate, a sentence in one of Leibniz's many writings to the effect that reality \emph{is} variety.
-Whether he made the statement or not, it is consistent with his view that any contingent thing is defined by a collection of attributes, which necessarily must be distinguishable
-at some epistemic level.
+The use of Shannon entropy $-\sum p_i \log p_i$ is well established in very different fields of knowledge such as computer science[cite], physics[cite] and ecology[cite].
+It is a fundamental element in information theory since Shannon developed his mathematical theory of communication.
+It has been used extensively within statistical mechanics to link the microscopic description of a system to its macroscopic thermodynamic one.
+It has been used in ecology to characterize biodiversity and calculate species distributions under given constraints.
+Entropy maximization has been shown to be a powerful tool in all these settings.
+
+While the techniques are well developed, it is puzzling that this concept can be applied in such different contexts, given that it assumes different features in each one.
+For example, in information theory the Shannon entropy for a book will change whether we choose to encode words or letters, which means the quantity does not have a fully objective feature: it depends on an arbitrary choice.
+This is not problematic for an information theoretic quantity, but it is for a physical quantity.
+In statistical mechanics the $p_i$ represent probabilities, which means the Shannon entropy is a property of an ensemble of physical systems.
+This leads some to claim that the Shannon entropy is epistemic in nature and represents ignorance.
+[Add references] In ecology, however, the $p_i$ represent the fraction of individuals belonging to the $i$th species.
+There is no probabilistic ensemble to which one can associate a degree of ignorance.
+Finally, thermodynamic entropy is associated to the notion of states and irreversible time evolution, which are not well defined concepts for the other fields.
+The same expression, then, seems to have acquire different flavors and it is not clear why a single concept would apply in all these cases.
+
+To make matter worse, it is fairly typical to describe entropy with notions such as uncertainty[cite], knowledge[cite], lack of knowledge[cite], disorder[cite].
+While some of these characterizations are useful in some contexts to give an intuitive picture, they do not seem to always work.
+A common objection to the characterization as disorder, for example, is that a uniform distribution corresponds to a homogeneous system, which one would expect to be maximally ordered; yet the distribution maximizes the Shannon entropy, which would indicate maximal disorder.
+Moreover, these notions remain vague in the sense that the formula is not formally derived by those concepts and therefore it is not clear why that particular indicator would be an appropriate choice.
+We are left to wonder: is there really some single underlying concept, that, when applied appropriately in the different fields, reduces to the correct notion of entropy?
+Or is it just a big coincidence?
+
+We propose that there is indeed a characterization of Shannon's formula that is both more general and more appropriate; that is independent on whether it is applied to sources of messages, ensembles of microstates or animal populations.
+We will argue that the formula is a measure of the \emph{variability of the values within a distribution of a given set of elements}.
+That is, given a distribution of some elements over some quantity, the Shannon entropy represents how varied, how different from each other, the values are.
+In fact, it is the only continuous and linear index that measures the degree of variability within a distribution, and those requirements lead naturally to same three conditions that Shannon set for his formula.
+In other words, the concept of variability with a distribution leads in a formal way to the correct expression.
+We will then show how this perspective naturally connects to information theory and statistical mechanics when applied in the respective contexts.
+
+As we believe this is the right characterization, we will therefore use the expression Shannon ``variability'', rather than ``entropy'' or ``information'', as it tells us exactly what the quantity describes.
+It is well known that when Shannon recognized the importance of his formula in communication theory he sought von Neumann's advice on the name he should give to the expression he had found.
+He was told to call it \emph{entropy}, first, because Boltzmann had used an identical formula to define entropy in statistical mechanics and, second, nobody understood entropy and that would provide protection from criticism [exact words to be found].
+We believe von Neumann's advice was not totally helpful and has given rise to confusion.
+We will reserve the word entropy for the thermodynamic entropy, the one dealing with physical states and irreversible processes.
+
+This characterization also explains why the Shannon variability has such a broad range of applicability.
+As nature presents itself in a multitude of forms and combinations, a fundamental task in science is to recognize which objects belong together and study how their different attributes are expressed in different ways or degrees.
+They may galaxies, animals or outcomes of the same process; we may study their shape, behavior or numerical quantities.
+But the idea of grouping them and try to understand the source of their variation is a common.\footnote{In fact, determining which collection of objects is significant and what attributes are relevant is significant part of the work.} Measuring and explaining this variability is an essential task.
+It is this very variety that is the inspiration for and the subject of scientific investigation.
+A universe without variety, where all objects were the same and remained the same, is barely conceivable and certainly it would leave no space for scientific investigation.
+
+As a final note, we want to stress that the goal of this work is not to present some new technical result.
+The goal is to put all details together in a coherent picture that gives them context and see precisely how they are related.
+This can help, especially the new practitioners, to avoid common misconceptions.
+Moreover, while some details presented here may be well known in a particular community, we find that it may be new to another.
+For example, we will show an example of Huffmann coding to illustrate the meaning of the numerical value associated with the Shannon variability.
+This is standard practice in computer science courses, yet it is not in statistical mechanics.
+Therefore restating all the crucial aspects in one place is beneficial because the connection between them is not always readily available.
+
+%https://www.researchgate.net/publication/259011656_A_tribute_to_Claude_Shannon_1916-2001_and_a_plea_for_more_rigorous_use_of_species_richness_species_diversity_and_the_'Shannon-Wiener'_Index
+
+%https://web.stanford.edu/~hastie/Papers/maxent_explained.pdf
+
+%https://www.mdpi.com/1099-4300/21/8/794/htm
+
+%Leibniz
+%We recall, though we have been unable to relocate, a sentence in one of Leibniz's many writings to the effect that reality \emph{is} variety.
+Whether he made the statement or not, it is consistent with his view that any contingent thing is defined by a collection of attributes, which necessarily must be distinguishable at some epistemic level.
 Paragraph 58 of Leibniz's \emph{Mondadology} asserts that the principle underlying the universe is that it should realize ``as much variety as possible, but with the greatest order possible''.
+[REF]
 
-Without committing ourselves to Leibniz's overarching principle, our aim in this paper is to draw attention to the specific nature of the variety that can be observed in the universe and the unique suitability of Shannon's entropy (or information) formula $-\sum_ip_i\,\textrm{ln}\,p_i$ for the characterization of the variety.
-Indeed, as we will show in the next section, opportunities for application of Shannon's formula are ubiquitous both on the earth and in the universe at large.
-
-We also wish to comment on the various interpretations and uses of Shannon's formula and attempt to answer this question: what does the formula represent?
-It is well known that when Shannon recognized the importance of his formula in communication theory he sought von Neumann's advice on the name he should give to the expression he had found.
-He was told to call it \emph{entropy}, first, because Botlzmann had used an identical formula to define entropy in statistical mechanics and, second, nobody understood entropy and that would provide protection from criticism [exact words to be found].
-We believe von Neumann's advice was not totally helpful and has given rise to confusion.
-Indeed, it is often claimed that Shannon's formula is epistemic in nature and represents ignorance.
-There certainly are situations in which that is true but equally there are others in which the use of the formula is very appropriate and no ignorance is involved at all.
-Examples will be given in the next section.
-
-This leads us to propose a different characterization of Shannon's formula that is both more general and more appropriate.
-Indeed, we will argue that the formula is a measure of the \emph{variability of the values within a distribution of a given set of elements} and we will use the expression Shannon ``variability'', rather than ``entropy'' or ``information''.
-Moreover, the formula has a very wide range of applicability because the universe contains many different kinds of things that, through having common attributes, can be recognized as belonging together despite the fact that their attributes are expressed to different degrees or in different ways.
-To give a first example, the girls in a school can differ in the length and colour of their hair.
-More examples will follow in the next section.
-In this paper we do not ask why the universe sustains the variety that it does.
-We take it as a given.
-
-We conclude this section with a word on terminology.
+%We conclude this section with a word on terminology.
 In mathematics, Cantor defined a set as ``a collection into a whole $M$ of definite and separate objects''.
@@
 mammals, birds, reptiles) and then later refine the mammals by species.
-What we would like to say is that the variability of the overall distribution $H_T = H_C + p_M H_M$ will be the variability over the classes $H_C$ plus the variability of the mammals $H_M$ weighted by the fraction of mammals $p_M$.
+In this case, we would like $H$ to combine linearly with respect to the probability.
+That is, we want the variability of the overall distribution $H_T = H_C + p_M H_M$ will be the variability over the classes $H_C$ plus the variability of the mammals $H_M$ weighted by the fraction of mammals $p_M$.
 This also makes the quantity linear when combining two independent distributions.

### 2019-10-29 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Reviewed intro with Christine

@@
 The use of Shannon entropy $-\sum p_i \log p_i$ is well established in very different fields of knowledge such as computer science[cite], physics[cite] and ecology[cite].
-It is a fundamental element in information theory since Shannon developed his mathematical theory of communication.
+Within computer science it is a fundamental element in information theory since Shannon developed his mathematical theory of communication.
 It has been used extensively within statistical mechanics to link the microscopic description of a system to its macroscopic thermodynamic one.
 It has been used in ecology to characterize biodiversity and calculate species distributions under given constraints.
-Entropy maximization has been shown to be a powerful tool in all these settings.
+Entropy maximization has been shown to be a powerful tool.
 
@@
 Finally, thermodynamic entropy is associated to the notion of states and irreversible time evolution, which are not well defined concepts for the other fields.
-The same expression, then, seems to have acquire different flavors and it is not clear why a single concept would apply in all these cases.
+The same expression, then, seems to have different traits and it is not clear why a single concept would apply in all these cases.
 
-To make matter worse, it is fairly typical to describe entropy with notions such as uncertainty[cite], knowledge[cite], lack of knowledge[cite], disorder[cite].
-While some of these characterizations are useful in some contexts to give an intuitive picture, they do not seem to always work.
+To make matters worse, it is fairly typical to describe entropy with notions such as uncertainty[cite], knowledge[cite], lack of knowledge[cite], disorder[cite].
+While some of these characterizations are useful in some contexts to give an intuitive picture, they do not always seem to work.
 A common objection to the characterization as disorder, for example, is that a uniform distribution corresponds to a homogeneous system, which one would expect to be maximally ordered; yet the distribution maximizes the Shannon entropy, which would indicate maximal disorder.
-Moreover, these notions remain vague in the sense that the formula is not formally derived by those concepts and therefore it is not clear why that particular indicator would be an appropriate choice.
+Moreover, these notions remain vague in the sense that the formula is not formally derived from those concepts, and therefore it is not clear why that particular indicator would be an appropriate choice.
 We are left to wonder: is there really some single underlying concept, that, when applied appropriately in the different fields, reduces to the correct notion of entropy?
@@
 
-We propose that there is indeed a characterization of Shannon's formula that is both more general and more appropriate; that is independent on whether it is applied to sources of messages, ensembles of microstates or animal populations.
+We propose that there is indeed a characterization of Shannon's formula that is both more general and more appropriate; that is independent of whether it is applied to sources of messages, ensembles of microstates or animal populations.
 We will argue that the formula is a measure of the \emph{variability of the values within a distribution of a given set of elements}.
 That is, given a distribution of some elements over some quantity, the Shannon entropy represents how varied, how different from each other, the values are.
-In fact, it is the only continuous and linear index that measures the degree of variability within a distribution, and those requirements lead naturally to same three conditions that Shannon set for his formula.
-In other words, the concept of variability with a distribution leads in a formal way to the correct expression.
-We will then show how this perspective naturally connects to information theory and statistical mechanics when applied in the respective contexts.
+In fact, it is the only continuous and linear index that measures the degree of variability within a distribution, and those requirements lead naturally to the same three conditions that Shannon set for his formula.\cite{Shannon} In other words, the concept of variability within a distribution leads in a formal way to the correct expression.
+In all cases where the formula is applied, in fact, we are using distributions as objects of our investigation and the variability within that distribution will be of interest.
+Yet, the physical meaning of those distributions can be widely different: they may represent the fractional distribution of animals according to their species; they may represent probability distributions of different outcomes; they may represent knowledge of certain facts as in Bayesian statistics.
+Therefore the variability itself, in the context of these different cases, will necessarily acquire a different meaning.
+In particular, we will see how this perspective naturally unfolds in the contexts of information theory and statistical mechanics.
 
-As we believe this is the right characterization, we will therefore use the expression Shannon ``variability'', rather than ``entropy'' or ``information'', as it tells us exactly what the quantity describes.
+As we believe this is the right characterization, we will therefore use the expression Shannon ``variability'', rather than ``entropy'', ``information'' or ``diversity index'', as it tells us exactly what the quantity describes.
 It is well known that when Shannon recognized the importance of his formula in communication theory he sought von Neumann's advice on the name he should give to the expression he had found.
@@
 We believe von Neumann's advice was not totally helpful and has given rise to confusion.
-We will reserve the word entropy for the thermodynamic entropy, the one dealing with physical states and irreversible processes.
+We will reserve the word ``entropy'' for the thermodynamic entropy, the one dealing with physical states and irreversible processes.
 
@@
 As nature presents itself in a multitude of forms and combinations, a fundamental task in science is to recognize which objects belong together and study how their different attributes are expressed in different ways or degrees.
-They may galaxies, animals or outcomes of the same process; we may study their shape, behavior or numerical quantities.
-But the idea of grouping them and try to understand the source of their variation is a common.\footnote{In fact, determining which collection of objects is significant and what attributes are relevant is significant part of the work.} Measuring and explaining this variability is an essential task.
+They may be galaxies, animals or outcomes of the same process; we may study their shape, behavior or numerical quantities.
+But the idea of grouping them and trying to understand the source of their variation is a common endeavor in science.\footnote{In fact, determining which collection of objects is significant and what attributes are relevant is a significant part of the work.} Measuring and explaining this variability is an essential task.
 It is this very variety that is the inspiration for and the subject of scientific investigation.
 A universe without variety, where all objects were the same and remained the same, is barely conceivable and certainly it would leave no space for scientific investigation.
+In the words of Leibniz, the principle underlying the universe is that it should realize ``as much variety as possible, but with the greatest order possible''.\footnote{MOVE TO BIBLIO Paragraph 58 of Leibniz's \emph{Mondadology}}
 
 As a final note, we want to stress that the goal of this work is not to present some new technical result.
-The goal is to put all details together in a coherent picture that gives them context and see precisely how they are related.
+The goal is instead to synthesize all details into a coherent picture that gives them context and a precise relationship to each other.
 This can help, especially the new practitioners, to avoid common misconceptions.
-Moreover, while some details presented here may be well known in a particular community, we find that it may be new to another.
-For example, we will show an example of Huffmann coding to illustrate the meaning of the numerical value associated with the Shannon variability.
+Moreover, while some details presented here may be well known in a particular community, we find that they may be new to another.
+For instance, we will show an example of Huffmann coding to illustrate the meaning of the numerical value associated with the Shannon variability.
 This is standard practice in computer science courses, yet it is not in statistical mechanics.
-Therefore restating all the crucial aspects in one place is beneficial because the connection between them is not always readily available.
+Therefore restating all the crucial aspects in one place is beneficial to make the connection between them readily available.
 
@@
 %https://www.mdpi.com/1099-4300/21/8/794/htm
+
+%https://www.sciencedirect.com/science/article/abs/pii/S0169534714001037
 
@@
 In this case, we would like $H$ to combine linearly with respect to the probability.
-That is, we want the variability of the overall distribution $H_T = H_C + p_M H_M$ will be the variability over the classes $H_C$ plus the variability of the mammals $H_M$ weighted by the fraction of mammals $p_M$.
-This also makes the quantity linear when combining two independent distributions.
+That is, we want the variability of the overall distribution $H_T = H_C + p_M H_M$ to be the variability over the classes $H_C$ plus the variability of the mammals $H_M$ weighted by the fraction of mammals $p_M$.
+This also makes the quantity additive when combining two independent distributions.
 

### 2019-11-08 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
After visiting Julian

@@
 
-%\usepackage{arxiv}
+\usepackage{amsmath}
 %\usepackage[usenames,dvipsnames]{color}
@@
 \author{
+	Gabriele Carcassi,
 	Julian Barbour and
-	Gabriele Carcassi
+	Christine A. Aidala 
 }
@@
 
+Now that we understand that the Shannon variability is measured by the number of bits required to identify an element from a distribution, it is instructive to derive the same expression from different considerations.
+Suppose we have a sequence of $N$ elements, say pets like in the previous example.
+These are taken from $I$ different cases: dogs, cats, fish and so on.
+Suppose $N_i$ are the of elements of each type, which means $\sum_{i \in I} N_i = N$.
+Then, given a particular instance of $N_i$, we have $W(N_i) = \frac{N!}{\prod_{i \in I} N_i!}$ possible ways to realize that sequence, which corresponds to all possible permutations.
+If all permutations are equally likely, then $\log W (N_i)$ represents the number of bits needed to identify one of the sequences.
+Assuming we use base $b$ for the logarithm, we take $N$ to be large and, using Stirling's approximation $\ln N!
+\approx N \ln N - N$, we find:
+\begin{equation}
+\begin{aligned}
+\log W(N_i) &= \log \frac{N!}{\prod_{i \in I} N_i!} \\
+&= \log N! - \sum_{i \in I} \log N_i! \\
+&\approx \frac{1}{\ln b} \left(N \ln N - N - \sum_{i \in I} N_i \ln N_i + \sum_{i \in I} N_i \right) \\
+&= \frac{1}{\ln b} \left(\sum_{i \in I} N_i \ln N - \sum_{i \in I} N_i \ln N_i \right) \\
+&= -\sum_{i \in I} N_i \left(\log N_i - \log N \right) \\
+&= - N \sum_{i \in I} \frac{N_i}{N} \log \frac{N_i}{N} \\
+&= - N \sum_{i \in I} p_i \log p_i \\
+&= N H(p_i)
+\end{aligned}
+\end{equation}
+
+The result should not be surprising: it simply tells us that encoding a sequence of $N$ elements is the same as encoding $N$ elements one at a time.
+
 It is important, at this point, to understand that the technical use of the term information in information theory does not equate to the normal use of the term which refers to knowledge, intelligible data.
@@
 There is no knowledge per se, just symbol manipulation that may represent different concepts in different contexts.
-
 The term information entropy, then, is misleading for two reasons.
@@
 
-So why is it so successfully used in statistical mechanics? Why does it seem to play a fundamental role even for systems far from equilibrium?[CITE]
+This should suggest that the success of Shannon variability and its use in both information theory and physics does not warrant a fundamental change of perspective in what constitutes a physical object.
+It is true that any physical process can be used to process information, when properly encoded.
+It is also true that scientific theories, in the end, are models that can capture only the aspects of nature that can be tested experimentally, the information extracted by the experiment, under suitable circumstances.
+Therefore the claim that information plays an important role in physical theories has a valid basis.[cite Wheeler, Landauer] But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
+The art of experimental science is neither contained in the mathematical description nor in the data collected.
+As information, in the information theoretic sense, requires that context to become intelligible, it cannot play a primary role.
+Therefore the claim that the universe itself is information does not follow.[cite Vedral, Davies]
+
+%Cite Vlatko Vedral: Decoding Reality: The Universe as Quantum Information. Information is "the fundamental building block of the Universe".
+
+%Cite Paul Davies Information and the Nature of Reality. Information "occupies the ontological basement"
+
+%Cite Wheeler It from bit.
+"It is wrong, moreover, to regard this or that  physical quantity as sitting "out there" with this or that numerical value it is absorbed." "The information thus solicited make physics and comes in bits"
+
+We have seen what the Shannon variability measures and why it is important in information theory, but it all assumed that the quantities we use for binning are discrete.
+In physics we are interested in quantities, like position and momentum, that are continuous.
+How does it work in that case?
+Wouldn't we need infinitely many bits to identify a continuous value from a continuous distribution?
+
+\section{Continuous distributions}
+
+A standard quick and dirty way to extend distributions from discrete variables to continuous variables is to substitute weights with densities and the sums with integrals.
+Therefore, instead of having a discrete normalize distribution $\sum_i p_i = 1$, we have a continuous normalized distribution $\int \rho(x)dx=1$.
+While simply changing $- \sum_i p_i \log p_i$ to $- \int \rho(x) \log \rho(x) dx$ works in most cases, it is easy to get confused because the two expressions have significantly different properties.
+
+In the discrete case, the Shannon variability is always positive. In the continuous case, the Shannon variability can be negative. For example, consider a uniform distribution over a line:
+\begin{equation}
+\rho(x)=
+\begin{cases}
+0 & x < a\\
+\dfrac{1}{b-a} & a \leq x \leq b\\
+0 & b < x
+\end{cases}
+\end{equation}
+We have $H(\rho) = - \int_a^b \frac{1}{b-a} \log \frac{1}{b-a} dx = \log (b-a)$. If $b-a < 1$ the Shannon variability will be negative. What does it mean to have negative number of bits?
+
+When considering continuous distributions, we have to remember that the densities are given per unit.
+That is, $\rho(x)$ has dimensions $[\frac{1}{x}]$.
+The variability will be also given relative to the unit.
+For example, a uniform distribution over a unit interval will correspond to zero variability.
+A uniform distribution over a two unit interval will correspond to one bit, since we will need one bit to narrow the variability back to one unit.
+A distribution over half a unit interval will correspond to minus one bit, since we would need to ``lose'' one bit of information to widen the variability back to one unit.
+In other words, the variability is measured by the number of bits needed to identify an element up to one unit.
+
+This means that infinitesimally narrow distributions, like delta functions, do not work well with the Shannon variability as they would give minus infinite entropy.
+If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which have a very small but still finite support (i.e.
+region where the function is non-zero).
+We will assume all our continuous distributions are like that.
+
+This brings up a second problem: what happens if we changed unit? If we changed the reference, we expect that the entropy will change accordingly. In fact:
+\begin{equation}
+\begin{aligned}
+H(\rho(y)) &= - \int \rho(y) \log \rho(y) dy \\
+&= - \int \rho(x) \left|\frac{\partial y}{\partial x}\right| \log \left(\rho(x)\left|\frac{\partial y}{\partial x}\right| \right) \left|\frac{\partial x}{\partial y}\right| dx \\
+&= - \int \rho(x)  \log \left(\rho(x)\left|\frac{\partial y}{\partial x}\right| \right)  dx \\
+&= - \int \rho(x)  \log \rho(x)  dx - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
+&= H(\rho(y)) - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
+\end{aligned}
+\end{equation}
+In general, the Shannon variability over a continuous variable is not invariant under coordinate transformations.
+Under translations, the Jacobian $\left|\frac{\partial y}{\partial x}\right|$ is unitary, so the Shannon variability does not change.
+If we stretch or shrink, if we change scale, the Shannon variability changes to measure the variability at the new scale.
+If the change is non-linear, the variability over different ranges will be counted differently.
+
+This should be particularly perplexing if we want to use this quantity in a physical setting.
+If we want $\rho(x)$ to represent the state of a macroscopic system and $H(\rho(x))$ to represent entropy, a state variable, how can $H$ change value if I merely expressed $\rho$ over different coordinates?
+How can maximization of entropy be meaningful if it yields different results depending on the coordinate system?
+Does that mean that we can use the Shannon formula only on distributions over discrete values?
+
+In physics, there is one thing that makes the Shannon variability work: phase-space.
+It is the only type of manifold where entropy can be defined in a way that is coordinate invariant.
+What happens in phase space is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components.
+Therefore the areas $d\hat{q}^i d\hat{p}_i = dq^j dp_j$ remain the same.
+The Jacobian determiants are unitary and the Shannon variability is invariant.
+
+The use of phase-space in statistical mechanics is of paramount importance: over this space both the density $\rho(q^i, p
+_i)$ and its variability $H(\rho(q^i, p
+_i))$ are invariant under arbitrary transformations of $q^i$, including non-linear ones.
+This is not just a coincidence or convenience: it is essential if we want to give the density at each point of phase space a physically objective character.
+
+The importance of phase space is often attributed to the Liouville's theorem, that states that areas are conserved under Hamiltonian evolution.
+While this is true, their invariance under coordinate transformations is more important as it is what makes them physically meaningful in the first place.
+In fact, how could we have conservation over time if we could not compare it in an observer independent way?
+This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\rho \log \rho$.
+If we use distributions over position and velocities or over momentum only, the expression would change for those particular variables in those coordinate systems.
+This can be, again, a source of confusion.\cite{Dunkel}
+
+We conclude noting that in quantum mechanics the Shannon variability, in the form of the Von Neumann entropy, is coordinate invariant as well.
+That is, given a density matrix $\rho$ then $S_{VN} = \textrm{tr}(\rho \log \rho)$ is independent on the basis in which it is calculated.
+
+Now that we have seen how the Shannon variability works in the continuous case and the importance of phase space, we are ready to see how it can be used in statistical mechanics and how it relates to the Boltzman, Gibbs and Von Neumann entropies.
 
@@
 As we saw in the previous sections, Shannon variability is a very general concept that applies to any type of distribution, no matter what the problem domain.
-It does not require us to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamic and statistical mechanics which requires them.\footnote{For this reason it is unfortunate that it was given the same name.} It is only when variability is appropriately applied to this particular setting that, though not apparent at first, a connection can be made.
-To understand this link we need to answer two basic questions.
-First, what does the Shannon variability represent in the context of statistical mechanics and thermodynamics?
-In particular, why can it be treated as a state variable and what property of the system is it describing?
-Second, why does the Shannon variability satisfy the same properties as the entropy?
-Namely, why does it increase during irreversible evolution, why is it maximized at equilibrium and why is it additive for independent systems?
+It does not require us to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamic and statistical mechanics which requires them.
+Yet it can, and is, applied successfully in statistical mechanics even in systems far from equilibrium.[cite] How does this work?
 
-The first question is straightforward to answer.
-The macroscopic variables of a system (e.g.
-volume, pressure, energy) do not give a full description of its parts.
-If we fix a macrostate and then measure the state of all its microscopic constituents, we will not get the same answer every time.
-This is true in the case of non-equilibrium (e.g.~we suddenly compress a gas by pushing a piston) and of equilibrium (e.g.~we give enough time to equilibrate).
-The choice of a macrostate, then, identifies a set of possible microstates that are compatible with it.
-It is then valid to ask the question: what is the variability of microstates one can find within a given macrostate?
-This is answered by the Shannon variability, which gives us the number of bits required to identify a microstate given the macrostate.
-Since the macrostates fully define the distribution of possible microstates, this distribution does not depend on the particular way the macrostate came to pass.
-It is path independent: it is a state variable.
-In short, as a macrostate is a distribution over microstates, it will admit as a state variable the Shannon variability, regardless of whether we know the actual distribution and can calculate its value.
+As statistical mechanics aims to describe the collective behavior of a large number of physical systems (i.e.
+particles), the Shannon variability can be applied to their distribution and will correspond to the entropy of the system.
+Yet, we have to understand that there are two distinct ways to apply the concept simply because there are two types of distributions one can be interested in.
+We may consider the state of the whole system at a given time, and be interested in the distribution of the particles over all possible particle states.
+This will lead to the Boltzman entropy.
+We may consider a statistical ensemble, which is a large collection of independent copies of the system found in different states, and consider the distribution of these different instances over the states of the whole system.
+This will lead to the Gibbs or the Von Neumann entropy depending on whether the system is classical or quantum.
+Both these types of distributions are used and are therefore of interest, so we will look at both and clarify the potential sources of confusion.
 
-Note, though, that the thermodynamic entropy is defined differently.
-It is defined over a succession of equilibrium states by noting that the integral of the heat exchanged divided by the temperature is path independent.
-One then chooses a particular state $A$, assigns it an arbitrary entropy value $S_A$ and the entropy for all other states is given by the integral $S(B) = S_A + \int_A^B \frac{dQ}{T}$.
-At this point, it is not apparent why these two quantities defined independently would be related.
-What we can say, though, is that all quantities that increase over irreversible processes, are maximized at equilibrium and that are additive for independent systems are precisely the quantities that are linear functions of entropy (see \cite{Giles}).
-A linear function of entropy, though, simply means a redefinition of $S_A$ and a change of units for either $dQ$ or $T$.
-Therefore if we can show that the Shannon variability possesses those three properties than it is equivalent to the entropy up to a change in units, which would answer our second question.
+Suppose we have a large number $N$ of particles taken from a normalized distribution $\rho(q^i, p_i)$.
+The space is the six dimensional phase-space for a single particle, sometimes called $\mu$-space.
+As we assume $N$ large, we can think of $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$ as the number of particles that are have within a region $\Sigma$ of phase space.
+That is, $\rho$ does not represent a probability distribution but an actual physical distribution that tells us the state of the whole system at one instant of time.
+The Boltzmann entropy is given by $S_B = k_B \log W$, where $W$ will correspond to the different ways that the $N$ particles can be arranged while still satisfying the distribution.
+If the space were discrete, the computation of the permutations would be straightforward.
+But how does this work in a continuous space?
 
-Suppose we have an irreversible process, which means we can start from many out-of-equilibrium states which end in a single equilibrium state.
-During the whole process, some of the macrostate variables are going to stay constant as they characterize the constraints to which the system is subjected during the evolution and will also identify the final equilibrium.
-For example, if the walls are fixed and insulated, the gas will have a set volume and average energy, which will lead to a unique equilibrium with a given temperature and pressure.\footnote{It can also happen that the irreversible process does not have a single possible equilibrium, as in the case of spontaneous symmetry breaking.
-New variables will identify the equilibrium which we can think of as new constraints appearing once the initial symmetry is broken, thus reducing to the previous case.} As different out-of-equilibrium macrostates merge during the evolution, they merge the possible microstates that lead to the same macrostate.
-The variability of microstates within the initial out-of-equilibrium macrostate, then, must be smaller than the variability of the microstates within the final equilibrium state.
-Irreversible processes increase the Shannon variability within the macrostate, which corresponds to a greater uniformity of the macrostate itself since uniform distributions are the ones that maximize the variability.
+As we said before, we treat continuous variables by comparing to a finite unit.
+We can pick a unit of phase space small enough such that the density $\rho$ is constant over cells of that size.
+We express $\rho$ as the number of particles within the chosen unit and divide phase space into cells.
+Now we can calculate all the possible permutations of the particles within the different cells and, in these circumstances, we will find that $\log W = N H(\rho)$.
+That is, the number of permutations at that level of precision will be equivalent to the number of particles times the variability of the distribution at that level of precision.
+In other words, the Boltzmann entropy reduces to the Shannon entropy.
 
-To understand why this is maximized at equilibrium, suppose it were not.
-That is, suppose there were another macrostate which satisfied the same constraints, but for which the Shannon variability were greater.
-If we prepared such a state and let it equilibrate, it could not evolve to the equilibrium because that has lower Shannon variability.
-Its equilibrium, then, would be different from the previous one, but this cannot be since we said that those constraints would lead to a single equilibrium state.
-Therefore the Shannon entropy is maximized for an equilibrium state.
+The Boltzmann constant $k_B$ should not distract us: its role is simply to allow us to measure temperature in an appropriate unit.
+If one defines $T = \frac{\partial U}{\partial S}$, and measures temperature in Kelvin and energy in Joules, the above relationship forces us to measure entropy in Joules/Kelvin.
+Therefore the entropy cannot be expressed as a pure number.
+However, this is not the only possible definition.
+Instead of using $T$ as a primary thermodynamic variable, we can use $\beta = \frac{1}{k_B T}$.\footnote{There are a few reasons why one can argue this is a better choice, though ultimately is still a matter of choice.} In this case, one would define $\beta = \frac{\partial S}{\partial U}$, measure energy in Joules, $\beta$ in inverse Joules and entropy would be dimensionless.
+As one can do all the thermodynamic calculations using just $\beta$, the constant $k_B$ is really just set by the unit system.
 
-Now suppose we have two independent systems, like two volumes of gasses.
-As they are independent, picking the microstate of one system does not constrain the microstate of the other.
-That is, the distribution over microstates are statistically independent.
-This means that the joint distribution is the product of the marginal distributions and, in this case, the Shannon entropy is additive.
+This setting, distributions over single particle phase space, is the one used to derive the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the correct distribution of particles for an ideal gas.
+It provides a clear picture for what happens in simple cases: the equilibrium distribution is the one where the variability of the states of the particles is maximal.
+That is, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles.
+The information needed to identify a particle within the distribution at a given set of precision will increase during an irreversible process.
+The equilibrium is a statistical equilibrium, particles are moving around, but for one particle that moves in one direction, there is one that moves in another and the overall distribution remains the same.
+TODO This corresponds to the insight of Boltzmann...
 
-The Shannon variability, then, has all the right properties to establish a conceptually meaningful connection with the thermodynamic/statistical mechanics entropy.
-It is a state variable that represents the variability of microstates within a macrostate.
-It increases during irreversible processes, reaches the maximum at equilibrium and it is additive for independent systems.
-It is indeed an independent concept but when applied to macrostates yields a state variable with the right conceptual characteristics.
-There are still some mathematical details that need to be worked out to have a complete picture.
-These details don't change the overall intuitive picture, but they present yet again opportunities for confusion.
+This approach, though, will only work in the limit of a large number of indistinguishable particles that are independently distributed.
+If $N$ is not large, we are not guaranteed that another particle will move in the opposite direction.
+If there are correlations, the motions will not cancel out.
+More precisely, suppose we have a probability distribution $\hat{\rho}$ for $N$ particles.
+If they are independently distributed, then the joint probability $\hat{\rho}=\prod_{i}\rho_i$ is the product of the distribution for each particle.
+If they are indistinguishable, then $\rho_i=\rho$: each particle has the same distribution.
+If $N$ is large, then number of particles in one region $N_\Sigma$ is very close to the expectation value $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$.
+If these assumptions are not met, the fluctuations will become relevant and we need a statistical theory to take them into account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.} Moreover, a change of the number of particles or particles of different species cannot be captured by a single normalized distribution over single particle states.
 
-The first is the role of the Boltzmann constant $k_B$, which is a conversion factor between the Shannon variability $H = - \sum p_i \log(p_i)$ and the Gibbs entropy $S_G = - k_B p_i \sum \log(p_i)$.
-This conversion constant should not surprise us, since we only established a linear relationship between variability and entropy.
-The question is whether this conversion constant itself has a physical meaning, or it is also simply due to a choice of units.
-The standard relationship between temperature, internal energy and entropy is $T = \frac{\partial U}{\partial S}$.
-Historically, temperature and internal energy are the quantities that were defined first.
-If we measure temperature in Kelvin and energy in Joules, the above relationship forces us to measure entropy in Joules/Kelvin.
-If we changed units, $k_B$ would change accordingly.
+The more general setting, then, is the following.
+The macrostate is a probability distribution over all possible complete description of the system, the microstates.
+That is, we have a distribution over the $6N$ dimensional phase space of $N$ particle, sometimes referred to as $\Gamma$-space, where each point represents the position and momentum of $N$ particles.
+The Gibbs entropy is $S_G = -k_B \int \rho \log \rho (dq^idp_i)$ which corresponds to the Shannon variability of the distribution.
+The Gibbs entropy, then, is the variability of a microstate as it moves around within the macrostate.
+The macrostate of an equilibrium will fully be identified by a set of macroscopic variables, such as temperature, average energy, pression, and so on.
+Note that these may be quantities that are not defined on an individual microstate but only on the ensemble.
+The microscopic dynamics will be free to move around as long as those statistical quantities are preserved.
+The Gibbs entropy, then, tells us the variability of the microstate under the given constraints and, at equilibrium, we will find that variability to be maximal.
 
-However, this is not the only way to define the relationships between those quantities.
-Another way is to have $\beta = \frac{1}{k_B T} = \frac{\partial S}{\partial U}$.
-In this setting, $U$ is measured in Joules, $S$ in nats (which are dimensionless), $\beta$ is measured in inverse Joules.
-Temperature is still measured in Kelvin, but we do not regard it as a primary object.
-We find this approach, sometimes called entropy first, preferable for a number of reasons, especially in statistical mechanics.
-First, entropy is invariant under coordinate transformations and therefore it is better to keep it a pure number and not mix it with units of energy, which is not invariant under boost.\footnote{This is essential if one ever hopes to put statistical mechanics into a relativistic setting.} Second, the maximization of entropy at equilibrium guaranties us that $S(U)$ is a single valued function (i.e.
-the maximum given the value of $U$) and $\frac{\partial S}{\partial U}$ is therefore well defined.
-On the other hand, $U(S)$ is not in general single valued (e.g.
-in the Ising model).
-This means that temperature can be infinite where the curve inverts, where small changes of energy do not change the entropy, and can be negative in the other region.
-That is, the infinitely cold system is temperature zero, which cannot be achieved, as it warms up temperature rises until it becomes infinite, and as the system becomes even hotter the temperature is negative and increase toward zero, which would correspond to an infinitely hot system.
-On the other hand, $\beta$ is always finite and cannot be infinite.
-An infinitely cold system, then would correspond to positive infinite $\beta$, as the system warms up beta decreases, passes from zero, becomes negative and an infinitely hot system corresponds to minus infinite $\beta$: this is a much better behavior for a physically meaningful quantity.
-Third, when setting up entropy maximization, $\beta$ is quite naturally the Lagrange multiplier associated with energy conservation, therefore mathematically it has a much clearer role.
-Lastly, one can proceed in both thermodynamic and statistical mechanics with just $\beta$, and convert to temperature only when a value on the Kelvin scale is needed.
-For these reasons, it seems to us $\beta$ plays the more fundamental role and $k_B$ is the conversion coefficient to go from $\beta$ to temperature measured in Kelvin.
-Adding and removing the Boltzmann constant from the entropy expression, then, is equivalent to shift from an energy/temperature first approach to an entropy/$\beta$ first approach and should be not seen as physically significant.
-It is really just fixing a unit.
+There are a couple of problems in this picture that we have to be aware of.
+The first is that $\Gamma$-space automatically assumes that all particles are distinguishable.
+This leads to the widely known problem of overcounting which needs to be addressed in the standard way.
+The second problem is that, though the state of each particle is given by position and momentum, we should not think of them as literally pointlike.
+As we say, this would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon entropy.
+It is more appropriate, both mathematically and conceptually, to think of indistinguishable particles as identical peaked distributions, each contributing the same amount of Shannon entropy.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.} The position and momentum corresponds to the center of mass of the particle.
 
-The other point of possible confusion is that in statistical mechanics there are actually two distributions.
-A single microstate is itself a distribution of particles over their particle states.
-There are actually $N$ particles existing at the same time, each with its own particle state.
-And a macrostate is a probability distribution over microstates.
-It is important to distinguish between these two distributions, especially because a microstate can be represented in two ways.
+We note that some authors choose to interpret the probability distribution not as representing the frequency over a large amount of independent trials, but as the knowledge one has about the system.
+This would make the entropy an observer dependent quantity, instead of an objective notion.\footnote{Jaynes called it the anthropomorphic nature of entropy.} The issue is that, experimentally, the fluctuations we observe are not observer dependent.
+We believe the confusion comes from the, correct, realization that the system by itself is not enough to determine the thermodynamic variables that are needed to reach the equilibrium.
+In fact, Jaynes points out:
+\begin{quote}
+Consider, for example, a crystal of Rochelle salt.
+For one set of experiments on it, we work with temperature, pressure, and volume.
+The entropy can be expressed as some function $S_e(T,P)$.
+For another set of experiments on the same crystal, we work with temperature, the component $e_xy$ of the strain tensor, and the component $P_z$ of electric polarization; the entropy as found in these experiments is a function $S_e(T,e_xy,P_z)$.
+It is clearly  meaningless to ask, ``What is the entropy of the crystal?''  unless we first specify the set of parameters which define its thermodynamic state.
+\end{quote}
+This is a particular case of what we stated in the previous sections: a distribution, and therefore its variability, is contingent upon some arbitrary choices.
+But, again, once those choices are made, the distribution is objective.
+The choice of the system, in statistical mechanics and thermodynamics, is enough to determine the state space for the microstates, but not enough to determine the set of macrostates that correspond to equilibria.
+We have to specify the process and the constraints that that process puts on the system.
+That is, under a different choice of process and constraints the microstates will fluctuate in different ways since we have changed the dynamics of the system.
+The question, then, is whether the system plus the process plus the set of constraints are enough to determine the probability, or we also need to mix in human knowledge.
+What we find is that they are sufficient, so the probability distribution does not depend on human knowledge.
 
-The first way is as a distribution over a single copy of phase space, sometimes referred to as $\mu$-space.
-In this case the distribution represents the fraction of particles that are in that state.
-The Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$ is expressed in this space.
-On $\mu$-space, the Shannon variability $-\int\rho \log \rho$ does not coincide with the thermodynamic entropy for a simple reason: it does not depend linearly on the number of particles.
-It is not an extensive quantity.
-The reason is straightforward: the Shannon entropy works on normalized distributions and gives us the variability of a single element.
-In this case, the system is composed by $N$ elements, not one.
-One can use the expression $-N\int\rho \log \rho$ as this represents the variability of $N$ elements.
+The case where the microstate is a quantum state is formally similar to the classical case.
+Instead of a distribution over the $N$ particle phase space, we have a distribution over the Hilbert space for the quantum system which is represented by a density matrix operator $\rho$.
+The Von Neumann entropy is given by $S_{VN} = \textrm{tr}(\rho \log \rho)$.
+This, expanded in a basis, becomes $S_{VN} = - \sum_i \rho_i \log \rho_i$ or $S_{VN} = - \int \rho(x) \log \rho(x)dx$ depending on whether the spectrum is discrete or continuous.
 
-We want to stress that a distribution on $\mu$-space can only either describe a single microstate where $N$ is very large and therefore the distribution can be assumed to be continuous, or a probability distribution for $N$ identically and independently prepared particles.
-A macrostate cannot be described by a single distribution on $\mu$-space as this cannot capture correlations.
-One would need to work with a probability distribution (i.e.
-the weights represents probabilities) over the space of particle distributions (i.e.
-the weights represent fractions of total particles) over $\mu$-space or one would obtain erroneous results.\footnote{Unfortunately this happens, for example, in [Jaynes] where he creates the distribution over $\mu$-space by averaging over all microstates.
-This essentially blends together the probabilities and particle counts losing most of the actual statistical information.
-Naturally, the entropy thus computed is wrong and found to not work.}
-
-Alternatively, a microstate can be represented as a point over $N$ copies of phase space, sometimes referred to as $\Gamma$-space.
-It is on this space that the Gibbs entropy is defined and where thermodynamic entropy happens to coincide with the Shannon entropy.
-Since this is a probability distribution, the microstate corresponds to a single element within the distribution defined by the macrostate.
-Yet, one has to be careful of two things.
-The first is that permutations of particles within $\Gamma$-space lead to different points which should not lead to different microstates.
-This leads to the widely known problem of overcounting which needs to be addressed in the standard way.
-The second problem is that we should not think of the particles as literally pointlike: they would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon entropy.
-It is more appropriate, both mathematically and conceptually, to think of indistinguishable particles as identical peaked distributions over $\mu$-space, each contributing the same amount of Shannon entropy.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.}
-
-The last point of confusion is the invariance of entropy under coordinate transformations.
-The use of phase-space in statistical mechanics is of paramount importance for one reason and one reason only: it is the only type of manifold where entropy can be defined in a way that is coordinate invariant.
-The issue is that we do not define fractions/probabilities over continuous variables: we defined densities.
-For a discrete case, we have a finite amount, like 3\%.
-For a continuous case we have an infinitesimal contribution, like 3\% per meter.
-If we switch labels in the discrete case, 3\% remains 3\%.
-If we switch to kilometers in the continuous case, we have 3000\% per kilometer.
-The continuous version of the Shannon entropy is affected by a coordinate transformation by the integral of the logarithm of the Jacobian determinant.
-Any transformation where the Jacobian determinant is not unitary, where the volumes are not preserved, does not preserve the Shannon entropy.
-This is not a problem for the Shannon entropy in general.
-It is a problem if we want to give the Shannon entropy any physical meaning.
-
-What happens in phase space is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components.
-Therefore the areas $d\hat{q}^i d\hat{p}_i = dq^j dp_j$ remain the same.
-The Jacobian determiants are unitary and the Shannon entropy is invariant.
-This brings us to another point of confusion: it is often argued that phase space areas are important because they are conserved under Hamiltonian evolution.
-While they are indeed conserved, that is not the reason why they are important.
-It is their invariance under coordinate transformation that makes them physically significant.
-In fact, conservation over time would not even be well defined if the quantity was not invariant and therefore have different values depending on the observer.
-
-This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\rho \log \rho$.
-If we use distributions over position and velocities or over momentum only, the expression would change for those particular variables in those coordinate systems.
-This can be, again, a source of confusion.\cite{Dunkel}
-
-The last issue is the relationship with the Boltzman entropy $S_B= k_B \log W$.
-This is actually captured by another standard way to derive the Shannon variability.
-Suppose we have a sequence $N$ elements, say twenty balls.
-There are $I$ possible types of elements, say three possible colors: red, green and blue.
-Suppose $N_i$ are the elements of each possible types, which means $\sum_{i \in I} N_i = N$.
-Then, given a set of $N_i$, we have $W(N_i) = \frac{N!}{\prod_{i \in I} N_i}$ possible ways to realize that sequence which corresponds to all possible permutations.
-If all permutations are equally likely, then $\log W$ represents the number of bits needed to identify one of the sequences.
-Since there are $N$ elements in the sequence, $\frac{1}{N} \log W$ represents the average number of bits needed to identify one element.
-We now take the limit for large $N$, that is we ask what is the average number of bits needed to identify one element from an infinite sequence given by the weights $p_i = \frac{N_i}{N}$.
-We find $\frac{1}{N} \log \frac{N!}{\prod_{i \in I} N_i} \approx - \sum p_i \log p_i$.
-In other words, the Boltzman entropy reduces to the Shannon entropy when counting permutations within large distributions.
-
-These technical details are important to understand the fine details of the links between the different entropies, they do not alter, and should not distract us, from the main conceptual point.
-The Shannon variability, the thermodynamic entropy, the Boltzmann entropy and the Gibbs entropy, though describing different things, are conceptually related.
-The variability of the state of the parts within the state of the whole is what the Shannon variability quantifies if applied to the macrostates of thermodynamics and the microstates of statistical mechanics.
-This variability increases during irreversible evolution, more microstates converge to the same macrostate until equilibrium is reached, where the variability is maximized.
-The variability of one system does not affect the variability of an independent one and therefore it simply sums.
-This is not a coincidence: this is what thermodynamic entropy ultimately is about.
-
-\section{Comments on Boltzmann's $H$ Theorem}
-
-[This is for you Gabriele].
-In Boltzmann's 1872 paper, he introduced the function that later was called the $H$ function, though it seems he meant to call it [upper case] $\eta$ and it was printed as $E$ (which somehow later became $H$).
-In the paper Boltzmann describes (in $\mu$-space) a gas of monatomic molecules that has already become spatially uniform and also has the same distribution wrt the particles' kinetic energies $x$ at each space point.
-He therefore considers a spatially constant distribution function $f(x,t)$ and wants to show that whatever its initial value it will tend in accordance with the Boltzmann equation (which he has previously derived and I won't give here) to the Maxwell distribution $f(x,t)=C\sqrt xe^{-hx}$ as $t\rightarrow\infty$.
-He defines his function $E$ as follows:
-\begin{equation}
-E=\int_0^\infty f(x,t)\left\{\textrm{log}\left [{f(x,t)\over \sqrt x}\right ]-1\right\}\textrm dx\label{hfun}
-\end{equation}
-and aims to prove that it cannot increase and attains its least value when $f(x,t)$ becomes the Maxwell distribution.
-Boltzmann does not give the argument that leads him to the function (\ref{hfun}), but somewhat later in the paper he notes that the $-1$ in the integrand merely reflects the fact that
-$$
-\int_0^\infty f(x,t)\textrm dx
-$$
-is the total number of molecules in unit volume, so that 
-\begin{equation}
-E=\int_0^\infty f(x,t)\textrm{log}\left [{f(x,t)\over \sqrt x}\right ]\textrm dx\label{hfunb}
-\end{equation}
-also cannot increase.
-
-Boltzmann does not say how he arrived at the function (\ref{hfun}), but its form is obviously dictated by the form of the Maxwell distribution.
-I'm terrible at equations, but I assume the $\sqrt x$ in the denominator of the integrand in (\ref{hfunb}) is because the usual Shannon expression is `skewed' by the subsidiary Lagrangian multiplier that fixes the total energy to a definite value.
-Later in the paper, when considering polyatomic molecules, he arrives by means of arguments I have not yet attempted to understand at the familiar $f\, \textrm{log}\,f$ form.
-
-In his very concise 1860 derivation of his energy distribution function, Maxwell commented that
-
-\begin{quote}\small
-
-It appears from this proposition that the velocities are distributed among the particles according  to the same law as the errors are distributed among the observations in the theory of the ``method of least squares''.
-
-\end{quote}\normalsize
-
-In his much more substantial 1867 paper, Maxwell noted that his 1860 derivation could be seen as `precarious' and gave an alternative proof.
-However, this does not eliminate the mathematical identity of his distribution with the error distribution.
-Would you agree that this (somewhat indirectly) establishes another example of variablity within a distribution of things [which I think we might use instead of `elements'; `members' (of a set) is another word we might consider]?
+The different entropies in statistical mechanics, then, all have a tight link to the Shannon variability.
+The Boltzmann entropy corresponds to the variability of a particle within a given microstate, provided that there are a large number of independent and indistinguishable particles.
+The Gibbs entropy corresponds to the variability of a classical microstate as constrained by the macrostate.
+The Von Neumann entropy corresponds to the quantum case.
+The characterization we gave to the Shannon formula, then, is readily applicable to statistical mechanics in a natural way.
+We find it more precise than the common characterizations, such as disorder or lack of knowledge, and, once internalized, leads to less confusion.
 

### 2019-11-09 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Wokring with Christine

@@
+%\documentclass[prb,preprint]{revtex4-1}
+
 \documentclass{article}
-\textheight 8.9 in \textwidth 6.5 in \oddsidemargin -10pt \topmargin
--30pt
+\textheight 8.9 in \textwidth 6.5 in \oddsidemargin -10pt \topmargin -30pt
 
@@
 
-\tableofcontents
+%\tableofcontents
 
@@
 While the techniques are well developed, it is puzzling that this concept can be applied in such different contexts, given that it assumes different features in each one.
-For example, in information theory the Shannon entropy for a book will change whether we choose to encode words or letters, which means the quantity does not have a fully objective feature: it depends on an arbitrary choice.
+For example, in information theory the Shannon entropy for a book will change whether we choose to encode words or letters, which means the quantity does not have a unique value: it depends on an arbitrary choice.
 This is not problematic for an information theoretic quantity, but it is for a physical quantity.
@@
 There is no probabilistic ensemble to which one can associate a degree of ignorance.
-Finally, thermodynamic entropy is associated to the notion of states and irreversible time evolution, which are not well defined concepts for the other fields.
+Finally, entropy in physics is associated to the notion of states and irreversible time evolution, which are not well defined concepts for the other fields.
 The same expression, then, seems to have different traits and it is not clear why a single concept would apply in all these cases.
@@
 We believe von Neumann's advice was not totally helpful and has given rise to confusion.
-We will reserve the word ``entropy'' for the thermodynamic entropy, the one dealing with physical states and irreversible processes.
+We will reserve the word ``entropy'' for the one dealing with physical states and irreversible processes.
 
@@
 Though this is a source of additional confusion, for the purpose of defining a measure of variability it does not change things conceptually.
-We will discuss the problem in connection to entropies in thermodynamics and statistical mechanics, since coordinate invariance is of paramount importance in physics.}
+We will discuss the problem in section \ref{sec_cont}.}
 
@@
 mammals, birds, reptiles) and then later refine the mammals by species.
-In this case, we would like $H$ to combine linearly with respect to the probability.
+In this case, we would like $H$ to combine linearly with respect to the weights.
 That is, we want the variability of the overall distribution $H_T = H_C + p_M H_M$ to be the variability over the classes $H_C$ plus the variability of the mammals $H_M$ weighted by the fraction of mammals $p_M$.
@@
 \end{enumerate}
-
 We note that these are exactly the requirements Shannon put forth for his expression\cite{Shannon}, and that these requirements lead to only one possible option: $H(p_i) = - \sum p_i \log(p_i)$.
 
-The Shannon entropy, which we will call Shannon variability, is a clear and important attribute of a given distribution: it quantifies the variability of the elements within the distribution.
-It quantifies the variety of values one finds within the distribution.
-It is true that the choices of the elements and binning are not fixed, it is true that the meaning of the weights depends on what the distribution is describing, and therefore that the precise meaning of this variability is context dependent.
+The Shannon variability is a clear and important attribute of a given distribution: it quantifies the variability of the elements within the distribution, the variety of values one finds.
+The precise meaning of this variability is context dependent, as the choices of the elements and binning are not fixed and the meaning of the weights depends on what the distribution is describing.
 But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope, a probability and so on.
-And like a real number represents a quantity, transcending what is physically quantified, the Shannon variability represents the variability of the elements\footnote{We again stress that out elements are physical and therefore possess some property beyond the bare `definite and separate' of Cantor's definition, which was perfectly suited to study numbers.
-Without these extra attributes, which are `washed out' by said definition, we would not be able to create a distribution in the first place.} in a distribution, transcending what the distribution physically signifies.
+And like a real number represents a quantity, transcending what is physically quantified, the Shannon variability represents the variability of the elements in a distribution, transcending what the distribution physically signifies.
 
@@
 
-Unfortunately, entropy in general is often associated with vague characterizations like the the two presented.
-It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending on the authors[CITATION NEEDED], which leads to confusion and misunderstanding.
-The characterization of the Shannon variability measure we have given, on the other hand, applies to all cases and leads naturally to the assumptions required to rederive it.
+Unfortunately, entropy in general is often associated with vague characterizations like the two presented.
+It is said to represent uncertainty, knowledge, lack of knowledge or disorder depending on the authors, which leads to confusion and misunderstanding.
+The characterization we have given of the Shannon variability measure, on the other hand, applies to all cases and leads naturally to the assumptions required to rederive it.
 Our characterization therefore is more fundamental.
@@
 
-This should suggest that the success of Shannon variability and its use in both information theory and physics does not warrant a fundamental change of perspective in what constitutes a physical object.
+This should suggest that the success of Shannon variability and its use in both information theory and physics does not warrant a fundamental change of perspective in what constitutes a physical object, as some physicists have claimed.
 It is true that any physical process can be used to process information, when properly encoded.
 It is also true that scientific theories, in the end, are models that can capture only the aspects of nature that can be tested experimentally, the information extracted by the experiment, under suitable circumstances.
-Therefore the claim that information plays an important role in physical theories has a valid basis.[cite Wheeler, Landauer] But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
+Therefore the claim that information plays an essential role in physical theories has a valid basis.[cite Wheeler, Landauer] But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
 The art of experimental science is neither contained in the mathematical description nor in the data collected.
@@
 
-We have seen what the Shannon variability measures and why it is important in information theory, but it all assumed that the quantities we use for binning are discrete.
-In physics we are interested in quantities, like position and momentum, that are continuous.
+We have seen what the Shannon variability measures and why it is important in information theory, but we have so far worked with discrete quantities.
+In physics we are also interested in continuous quantities, like position and momentum.
 How does it work in that case?
-Wouldn't we need infinitely many bits to identify a continuous value from a continuous distribution?
+Wouldn't we need infinitely many bits to identify a value from a continuous distribution?
 
-\section{Continuous distributions}
+\section{Continuous distributions}\label{sec_cont}
 
 A standard quick and dirty way to extend distributions from discrete variables to continuous variables is to substitute weights with densities and the sums with integrals.
-Therefore, instead of having a discrete normalize distribution $\sum_i p_i = 1$, we have a continuous normalized distribution $\int \rho(x)dx=1$.
+Therefore, instead of having a discrete normalized distribution $\sum_i p_i = 1$, we have a continuous normalized distribution $\int \rho(x)dx=1$.
 While simply changing $- \sum_i p_i \log p_i$ to $- \int \rho(x) \log \rho(x) dx$ works in most cases, it is easy to get confused because the two expressions have significantly different properties.
@@
 \end{equation}
-We have $H(\rho) = - \int_a^b \frac{1}{b-a} \log \frac{1}{b-a} dx = \log (b-a)$. If $b-a < 1$ the Shannon variability will be negative. What does it mean to have negative number of bits?
+We have $H(\rho) = - \int_a^b \frac{1}{b-a} \log \frac{1}{b-a} dx = \log (b-a)$. If $0 < b-a < 1$ the Shannon variability will be negative. What does it mean to have a negative number of bits?
 
@@
 
-This brings up a second problem: what happens if we changed unit? If we changed the reference, we expect that the entropy will change accordingly. In fact:
+This brings up a second problem: what happens if we change unit? If we change the reference, we expect that the entropy will change accordingly. In fact:
 \begin{equation}
@@
 &= - \int \rho(x)  \log \rho(x)  dx - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
-&= H(\rho(y)) - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
+&= H(\rho(x)) - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
 \end{aligned}
@@
 This should be particularly perplexing if we want to use this quantity in a physical setting.
-If we want $\rho(x)$ to represent the state of a macroscopic system and $H(\rho(x))$ to represent entropy, a state variable, how can $H$ change value if I merely expressed $\rho$ over different coordinates?
+If we want $\rho(x)$ to represent the state of a macroscopic system and $H(\rho(x))$ to represent entropy, a state variable, how can $H$ change value if we merely express $\rho$ over different coordinates?
 How can maximization of entropy be meaningful if it yields different results depending on the coordinate system?
@@
 In physics, there is one thing that makes the Shannon variability work: phase-space.
-It is the only type of manifold where entropy can be defined in a way that is coordinate invariant.
 What happens in phase space is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components.
@@
 
-The importance of phase space is often attributed to the Liouville's theorem, that states that areas are conserved under Hamiltonian evolution.
+The importance of phase space is often attributed to Liouville's theorem, that states that areas are conserved under Hamiltonian evolution.
 While this is true, their invariance under coordinate transformations is more important as it is what makes them physically meaningful in the first place.
 In fact, how could we have conservation over time if we could not compare it in an observer independent way?
-This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\rho \log \rho$.
+This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\int \rho \log \rho$.
 If we use distributions over position and velocities or over momentum only, the expression would change for those particular variables in those coordinate systems.
@@
 We conclude noting that in quantum mechanics the Shannon variability, in the form of the Von Neumann entropy, is coordinate invariant as well.
-That is, given a density matrix $\rho$ then $S_{VN} = \textrm{tr}(\rho \log \rho)$ is independent on the basis in which it is calculated.
+That is, given a density matrix $\rho$ then $S_{VN} = - \textrm{tr}(\rho \log \rho)$ is independent of the basis in which it is calculated.
 
@@
 As we saw in the previous sections, Shannon variability is a very general concept that applies to any type of distribution, no matter what the problem domain.
-It does not require us to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamic and statistical mechanics which requires them.
+It does not require us to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamics and statistical mechanics.
 Yet it can, and is, applied successfully in statistical mechanics even in systems far from equilibrium.[cite] How does this work?
@@
 The space is the six dimensional phase-space for a single particle, sometimes called $\mu$-space.
-As we assume $N$ large, we can think of $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$ as the number of particles that are have within a region $\Sigma$ of phase space.
+As we assume $N$ large, we can think of $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$ as the number of particles that are within a region $\Sigma$ of phase space.
 That is, $\rho$ does not represent a probability distribution but an actual physical distribution that tells us the state of the whole system at one instant of time.
@@
 That is, the number of permutations at that level of precision will be equivalent to the number of particles times the variability of the distribution at that level of precision.
-In other words, the Boltzmann entropy reduces to the Shannon entropy.
+In other words, the Boltzmann entropy reduces to $N$ times the Shannon variability of the single-particle distribution for a given microstate.
 
@@
 However, this is not the only possible definition.
-Instead of using $T$ as a primary thermodynamic variable, we can use $\beta = \frac{1}{k_B T}$.\footnote{There are a few reasons why one can argue this is a better choice, though ultimately is still a matter of choice.} In this case, one would define $\beta = \frac{\partial S}{\partial U}$, measure energy in Joules, $\beta$ in inverse Joules and entropy would be dimensionless.
+Instead of using $T$ as a primary thermodynamic variable, we can use $\beta = \frac{1}{k_B T}$.\footnote{There are a few reasons why one can argue this is a better choice, though ultimately it is still a matter of choice.} In this case, one would define $\beta = \frac{\partial S}{\partial U}$, measure energy in Joules, $\beta$ in inverse Joules and entropy would be dimensionless.
 As one can do all the thermodynamic calculations using just $\beta$, the constant $k_B$ is really just set by the unit system.
 
-This setting, distributions over single particle phase space, is the one used to derive the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the correct distribution of particles for an ideal gas.
+This setting, distributions over single particle phase space, is the one used to derive the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the distribution of particles for an ideal gas.
 It provides a clear picture for what happens in simple cases: the equilibrium distribution is the one where the variability of the states of the particles is maximal.
 That is, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles.
-The information needed to identify a particle within the distribution at a given set of precision will increase during an irreversible process.
-The equilibrium is a statistical equilibrium, particles are moving around, but for one particle that moves in one direction, there is one that moves in another and the overall distribution remains the same.
-TODO This corresponds to the insight of Boltzmann...
+The information needed to identify a particle within the distribution at a given precision will increase during an irreversible process.
+The equilibrium is a statistical equilibrium, particles are moving around, but for one particle that moves in one direction, there is another one that moves in the opposite and the overall distribution remains the same.\footnote{This corresponds to the insight developed by Boltzmann that led to the creation of statistical mechanics.}
 
 This approach, though, will only work in the limit of a large number of indistinguishable particles that are independently distributed.
-If $N$ is not large, we are not guaranteed that another particle will move in the opposite direction.
-If there are correlations, the motions will not cancel out.
+If $N$ is not large, there may not be another particle that moves in the opposite direction.
+If there are correlations, the motions may not cancel out.
 More precisely, suppose we have a probability distribution $\hat{\rho}$ for $N$ particles.
@@
 If they are indistinguishable, then $\rho_i=\rho$: each particle has the same distribution.
-If $N$ is large, then number of particles in one region $N_\Sigma$ is very close to the expectation value $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$.
-If these assumptions are not met, the fluctuations will become relevant and we need a statistical theory to take them into account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.} Moreover, a change of the number of particles or particles of different species cannot be captured by a single normalized distribution over single particle states.
+If $N$ is large, the number of particles in one region $N_\Sigma$ is very close to the expectation value $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$.
+If these assumptions are not met, we cannot break the joint distribution into single particle ones, the fluctuations may become relevant thus requiring a more general account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.} Moreover, a change of the number of particles or particles of different species cannot be captured by a single normalized distribution over single particle states.
 
 The more general setting, then, is the following.
-The macrostate is a probability distribution over all possible complete description of the system, the microstates.
-That is, we have a distribution over the $6N$ dimensional phase space of $N$ particle, sometimes referred to as $\Gamma$-space, where each point represents the position and momentum of $N$ particles.
-The Gibbs entropy is $S_G = -k_B \int \rho \log \rho (dq^idp_i)$ which corresponds to the Shannon variability of the distribution.
+The macrostate is a probability distribution over all possible complete descriptions of the system, the microstates.
+That is, we have a distribution over the $6N$ dimensional phase space of $N$ particles, sometimes referred to as $\Gamma$-space, where each point represents the position and momentum of $N$ particles.
+The Gibbs entropy is $S_G = -k_B \int \rho \log \rho \, dq^idp_i$, which corresponds to the Shannon variability of the microstate distribution for a given classical macrostate.
 The Gibbs entropy, then, is the variability of a microstate as it moves around within the macrostate.
-The macrostate of an equilibrium will fully be identified by a set of macroscopic variables, such as temperature, average energy, pression, and so on.
+The macrostate of an equilibrium will be fully identified by a set of macroscopic variables, such as temperature, average energy, pressure, and so on.
 Note that these may be quantities that are not defined on an individual microstate but only on the ensemble.
@@
 The first is that $\Gamma$-space automatically assumes that all particles are distinguishable.
-This leads to the widely known problem of overcounting which needs to be addressed in the standard way.
-The second problem is that, though the state of each particle is given by position and momentum, we should not think of them as literally pointlike.
-As we say, this would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon entropy.
+This leads to the widely known problem of overcounting which needs to be addressed in the standard way.[cite] The second problem is that, though the state of each particle is given by position and momentum, we should not think of them as literally pointlike.
+As we said, this would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon variability.
 It is more appropriate, both mathematically and conceptually, to think of indistinguishable particles as identical peaked distributions, each contributing the same amount of Shannon entropy.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.} The position and momentum corresponds to the center of mass of the particle.
@@
 The entropy can be expressed as some function $S_e(T,P)$.
-For another set of experiments on the same crystal, we work with temperature, the component $e_xy$ of the strain tensor, and the component $P_z$ of electric polarization; the entropy as found in these experiments is a function $S_e(T,e_xy,P_z)$.
+For another set of experiments on the same crystal, we work with temperature, the component $e_{xy}$ of the strain tensor, and the component $P_z$ of electric polarization; the entropy as found in these experiments is a function $S_e(T,e_{xy},P_z)$.
 It is clearly  meaningless to ask, ``What is the entropy of the crystal?''  unless we first specify the set of parameters which define its thermodynamic state.
@@
 That is, under a different choice of process and constraints the microstates will fluctuate in different ways since we have changed the dynamics of the system.
-The question, then, is whether the system plus the process plus the set of constraints are enough to determine the probability, or we also need to mix in human knowledge.
-What we find is that they are sufficient, so the probability distribution does not depend on human knowledge.
+The question, then, is whether the system plus the process plus the set of constraints are enough to determine the entropy, or if we also need to mix in human knowledge.
+What we find is that they are sufficient, so the entropy does not depend on human knowledge.
 
-The case where the microstate is a quantum state is formally similar to the classical case.
+The case of a quantum system is formally similar to the classical one.
 Instead of a distribution over the $N$ particle phase space, we have a distribution over the Hilbert space for the quantum system which is represented by a density matrix operator $\rho$.
-The Von Neumann entropy is given by $S_{VN} = \textrm{tr}(\rho \log \rho)$.
+The Von Neumann entropy is given by $S_{VN} = - \textrm{tr}(\rho \log \rho)$.
 This, expanded in a basis, becomes $S_{VN} = - \sum_i \rho_i \log \rho_i$ or $S_{VN} = - \int \rho(x) \log \rho(x)dx$ depending on whether the spectrum is discrete or continuous.
+The Von Neumann entropy corresponds to the Shannon variability of the microstate distribution for a given quantum macrostate.
 
 The different entropies in statistical mechanics, then, all have a tight link to the Shannon variability.
-The Boltzmann entropy corresponds to the variability of a particle within a given microstate, provided that there are a large number of independent and indistinguishable particles.
+The Boltzmann entropy corresponds to the variability of the state of a particle within a given microstate, provided that there are a large number of independent and indistinguishable particles.
 The Gibbs entropy corresponds to the variability of a classical microstate as constrained by the macrostate.
-The Von Neumann entropy corresponds to the quantum case.
+The Von Neumann entropy is similar but corresponds to the quantum case.
 The characterization we gave to the Shannon formula, then, is readily applicable to statistical mechanics in a natural way.
-We find it more precise than the common characterizations, such as disorder or lack of knowledge, and, once internalized, leads to less confusion.
+We find it more precise than the common characterizations, such as disorder or lack of knowledge, and, once internalized, it leads to less confusion.
 

### 2019-11-14 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
Preparing version for AJP

--- 2019-11-09 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
@@
-%\documentclass[prb,preprint]{revtex4-1}
+\documentclass[prb, twocolumn]{revtex4-1}
 
-\documentclass{article}
-\textheight 8.9 in \textwidth 6.5 in \oddsidemargin -10pt \topmargin -30pt
+%\documentclass{article}
+%\textheight 8.9 in \textwidth 6.5 in \oddsidemargin -10pt \topmargin -30pt
 
@@
 
+\begin{document}
+
 \title{Universal Characterization of Physical Variability by \\Shannon's $-\sum_ip_i\,\textrm{ln}\,p_i$ Formula}
 
-\author{
-	Gabriele Carcassi,
-	Julian Barbour and
-	Christine A. Aidala 
-}
+\author{Gabriele Carcassi}
+\author{Julian Barbour}
+\author{Christine A. Aidala}
 
-\begin{document}
+\begin{abstract}
+	The Shannon entropy, one of the cornerstones of information theory, is widely used in physics, particularly in statistical mechanics.
+Yet its characterization and connection to physics remains vague, leaving ample room for misconceptions and misunderstanding.
+In this paper we will show that the Shannon entropy can be fully understood as measuring the variability of the elements within a distribution: it characterizes how much variation can be found within a collection of objects.
+We will see that it is the only indicator that is continuous and linear, that quantifies the number of yes/no questions (i.e.
+bits) that are needed to identify an element within the distribution, and we will see how applying this concept to statistical mechanics in different ways lead to the Boltzman, Gibbs and Von Neumann entropies.
+\end{abstract}
 
@@
 
-\begin{abstract}
-TBD
-\end{abstract}
-
 \section{Introduction\label{int}}
 
-The use of Shannon entropy $-\sum p_i \log p_i$ is well established in very different fields of knowledge such as computer science[cite], physics[cite] and ecology[cite].
-Within computer science it is a fundamental element in information theory since Shannon developed his mathematical theory of communication.
-It has been used extensively within statistical mechanics to link the microscopic description of a system to its macroscopic thermodynamic one.
-It has been used in ecology to characterize biodiversity and calculate species distributions under given constraints.
-Entropy maximization has been shown to be a powerful tool.
+The Shannon entropy $-\sum p_i \log p_i$ is well established in such disparate fields as computer science[cite], physics[cite] and ecology[cite].
+It is most often introduced in physics through statistical mechanics, and associated to vague notions, such as uncertainty[cite], knowledge[cite], lack of knowledge[cite] and disorder[cite].
+These prevent a proper understanding of the concept and often lead to confusion and misunderstandings.
 
-While the techniques are well developed, it is puzzling that this concept can be applied in such different contexts, given that it assumes different features in each one.
-For example, in information theory the Shannon entropy for a book will change whether we choose to encode words or letters, which means the quantity does not have a unique value: it depends on an arbitrary choice.
-This is not problematic for an information theoretic quantity, but it is for a physical quantity.
-In statistical mechanics the $p_i$ represent probabilities, which means the Shannon entropy is a property of an ensemble of physical systems.
-This leads some to claim that the Shannon entropy is epistemic in nature and represents ignorance.
-[Add references] In ecology, however, the $p_i$ represent the fraction of individuals belonging to the $i$th species.
-There is no probabilistic ensemble to which one can associate a degree of ignorance.
-Finally, entropy in physics is associated to the notion of states and irreversible time evolution, which are not well defined concepts for the other fields.
-The same expression, then, seems to have different traits and it is not clear why a single concept would apply in all these cases.
+This paper aims to give a crisp characterization of the Shannon entropy that is intuitive and precise.
+The general idea is that it measures the variability of the elements within a distribution, which is a general concept applicable to many branches of science and it is independent of the notion of entropy of thermodynamics and statistical mechanics.
+Therefore we will use the term \emph{``Shannon variability''}, leaving ``entropy'' to the physical concept.
+We will show that the expression $- \sum_i p_i \log p_i$ is the only linear indicator of variability.
+It measures the number of questions one must ask to identify an element of the distribution, linking its use to information theory.
+We will study how the expression works over continuous variables, and show how phase-space is special as it leaves the Shannon variability invariant under change of coordinates.
+We will turn to statistical mechanics and apply Shannon variability in different ways to recover connections to the Boltzmann, Gibbs and Von Neumann entropies.
 
-To make matters worse, it is fairly typical to describe entropy with notions such as uncertainty[cite], knowledge[cite], lack of knowledge[cite], disorder[cite].
-While some of these characterizations are useful in some contexts to give an intuitive picture, they do not always seem to work.
-A common objection to the characterization as disorder, for example, is that a uniform distribution corresponds to a homogeneous system, which one would expect to be maximally ordered; yet the distribution maximizes the Shannon entropy, which would indicate maximal disorder.
-Moreover, these notions remain vague in the sense that the formula is not formally derived from those concepts, and therefore it is not clear why that particular indicator would be an appropriate choice.
-We are left to wonder: is there really some single underlying concept, that, when applied appropriately in the different fields, reduces to the correct notion of entropy?
-Or is it just a big coincidence?
-
-We propose that there is indeed a characterization of Shannon's formula that is both more general and more appropriate; that is independent of whether it is applied to sources of messages, ensembles of microstates or animal populations.
-We will argue that the formula is a measure of the \emph{variability of the values within a distribution of a given set of elements}.
-That is, given a distribution of some elements over some quantity, the Shannon entropy represents how varied, how different from each other, the values are.
-In fact, it is the only continuous and linear index that measures the degree of variability within a distribution, and those requirements lead naturally to the same three conditions that Shannon set for his formula.\cite{Shannon} In other words, the concept of variability within a distribution leads in a formal way to the correct expression.
-In all cases where the formula is applied, in fact, we are using distributions as objects of our investigation and the variability within that distribution will be of interest.
-Yet, the physical meaning of those distributions can be widely different: they may represent the fractional distribution of animals according to their species; they may represent probability distributions of different outcomes; they may represent knowledge of certain facts as in Bayesian statistics.
-Therefore the variability itself, in the context of these different cases, will necessarily acquire a different meaning.
-In particular, we will see how this perspective naturally unfolds in the contexts of information theory and statistical mechanics.
-
-As we believe this is the right characterization, we will therefore use the expression Shannon ``variability'', rather than ``entropy'', ``information'' or ``diversity index'', as it tells us exactly what the quantity describes.
-It is well known that when Shannon recognized the importance of his formula in communication theory he sought von Neumann's advice on the name he should give to the expression he had found.
-He was told to call it \emph{entropy}, first, because Boltzmann had used an identical formula to define entropy in statistical mechanics and, second, nobody understood entropy and that would provide protection from criticism [exact words to be found].
-We believe von Neumann's advice was not totally helpful and has given rise to confusion.
-We will reserve the word ``entropy'' for the one dealing with physical states and irreversible processes.
-
-This characterization also explains why the Shannon variability has such a broad range of applicability.
-As nature presents itself in a multitude of forms and combinations, a fundamental task in science is to recognize which objects belong together and study how their different attributes are expressed in different ways or degrees.
-They may be galaxies, animals or outcomes of the same process; we may study their shape, behavior or numerical quantities.
-But the idea of grouping them and trying to understand the source of their variation is a common endeavor in science.\footnote{In fact, determining which collection of objects is significant and what attributes are relevant is a significant part of the work.} Measuring and explaining this variability is an essential task.
-It is this very variety that is the inspiration for and the subject of scientific investigation.
-A universe without variety, where all objects were the same and remained the same, is barely conceivable and certainly it would leave no space for scientific investigation.
-In the words of Leibniz, the principle underlying the universe is that it should realize ``as much variety as possible, but with the greatest order possible''.\footnote{MOVE TO BIBLIO Paragraph 58 of Leibniz's \emph{Mondadology}}
-
-As a final note, we want to stress that the goal of this work is not to present some new technical result.
-The goal is instead to synthesize all details into a coherent picture that gives them context and a precise relationship to each other.
-This can help, especially the new practitioners, to avoid common misconceptions.
-Moreover, while some details presented here may be well known in a particular community, we find that they may be new to another.
-For instance, we will show an example of Huffmann coding to illustrate the meaning of the numerical value associated with the Shannon variability.
-This is standard practice in computer science courses, yet it is not in statistical mechanics.
-Therefore restating all the crucial aspects in one place is beneficial to make the connection between them readily available.
-
-%https://www.researchgate.net/publication/259011656_A_tribute_to_Claude_Shannon_1916-2001_and_a_plea_for_more_rigorous_use_of_species_richness_species_diversity_and_the_'Shannon-Wiener'_Index
-
-%https://web.stanford.edu/~hastie/Papers/maxent_explained.pdf
-
-%https://www.mdpi.com/1099-4300/21/8/794/htm
-
-%https://www.sciencedirect.com/science/article/abs/pii/S0169534714001037
-
-%Leibniz
-%We recall, though we have been unable to relocate, a sentence in one of Leibniz's many writings to the effect that reality \emph{is} variety.
-Whether he made the statement or not, it is consistent with his view that any contingent thing is defined by a collection of attributes, which necessarily must be distinguishable at some epistemic level.
-Paragraph 58 of Leibniz's \emph{Mondadology} asserts that the principle underlying the universe is that it should realize ``as much variety as possible, but with the greatest order possible''.
-[REF]
-
-%We conclude this section with a word on terminology.
-In mathematics, Cantor defined a set as ``a collection into a whole $M$ of definite and separate objects''.
-He called the objects \emph{elements}.
-We will also use the word `elements' to denote the things we consider but subject to the proviso that they have a distinguishing feature (or features) that allows us to say that \emph{they are of the same kind}.
-Thus, we consider sets of a restricted type.
-The point is this.
-To constitute a set in accordance with Cantor's definition, it is only necessary for its elements to be \emph{definite and separate} objects.
-This leaves total freedom as to how such objects in the universe are to be parcelled up into specific sets.
-For example, a set might consist of Mars and the word `go'.
-We think such latitude misses one of the most significant aspects of the universe, namely, that the things within it can be distinguished by attributes that are much more specific than the two required by Cantor.
-Our field of investigation goes beyond mere number, which washes out all attributes apart from the two, of being definite and separate, that allow us to count.
-They were the only attributes that Cantor needed; we need more.
+While some details presented here may be well known in a particular community, we find that they may be new to another.
+Therefore we assume little knowledge of the subject from the reader and include all details that we feel are important to properly understand the subject without confusion, including two standard derivations for the formula and examples that are common in computer science but rarely used in physics.
+We will briefly touch over a few of the common misconceptions.
 
@@
 \end{enumerate}
-We note that these are exactly the requirements Shannon put forth for his expression\cite{Shannon}, and that these requirements lead to only one possible option: $H(p_i) = - \sum p_i \log(p_i)$.
-
-The Shannon variability is a clear and important attribute of a given distribution: it quantifies the variability of the elements within the distribution, the variety of values one finds.
+These are the same requirements Shannon put forth for his expression\cite{Shannon}, from which he showed to be the only possible choice: $H(p_i) = - \sum p_i \log(p_i)$.
+This expression quantifies the variability of the elements within the distribution, the variety of values one finds.
 The precise meaning of this variability is context dependent, as the choices of the elements and binning are not fixed and the meaning of the weights depends on what the distribution is describing.
 But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope, a probability and so on.
-And like a real number represents a quantity, transcending what is physically quantified, the Shannon variability represents the variability of the elements in a distribution, transcending what the distribution physically signifies.
 
@@
 If we look at the words in the complete works of Shakespeare, what variability do we find in his vocabulary?
-All these questions are interesting, all these questions are insightful and the answers to all these questions are expressed through just one thing: the Shannon variability.
 
-If the Shannon expression is a measure of variability, how is it connected to information? What is it measuring and in what units?
+If the Shannon expression is a measure of variability, why is it connected to information? How is variability quantified and in what units?
 
@@
 
-We have seen that the Shannon variability measures the variability within a distribution but, in the end, it is a numeric quantity and we cannot expect to understand the concept if we do not understand what the number represents and in what units.
-
-We can gain insight with the following example.
+To understand what the numerical value represents, consider this example.
 Suppose we fix a distribution, say the animals in the Galap\'{a}gos islands binned by their respective species.
-Suppose we pick a specific animal from the set and you want to know its species.
+Suppose I pick a specific animal from the set and you want to know its species.
 Suppose the only way for you to get that information is to ask a series of questions with only two possible answers, yes or no.
@@
 Others, like ``is it an American Flamingo (Phoenicopterus ruber)?", would give us a lot of information in the positive case but little in the negative case.
-However, it is clear that there is always a minimum number of questions that must be asked to get to the answer.
-Given that there are more than two species, a single question cannot be enough.
-It should however be intuitively clear that to a greater variability within the distribution will correspond a greater number of questions you must ask.
+However, there has to be a minimum number of questions that must be asked to get to the answer.
+A single question, for example, cannot be enough given that there are more than two species.
+It should be intuitively clear that to a greater variability within the distribution will correspond a greater number of questions you must ask.
 That is exactly what the Shannon variability quantifies: the minimum average number of questions one has to ask to identify a value in the distribution.
@@
 We can also pick a non-integer base, like the natural base for logarithms, and we will have nats.
-This is why the Shannon variability is fundamental in information theory, because it quantifies how much information is needed to transfer a value from a distribution.
+This is why the Shannon variability is fundamental in information theory, because it quantifies how much information is needed to transfer a value picked from a distribution.
 
@@
 For example, suppose the population of pets in a country is as follows:
-\begin{itemize}
-\item dogs 27\%
-\item cats 48\%
-\item fish 10\%
-\item bird 8\%
-\item small mammal 4\%
-\item reptile 3\%
-\end{itemize}
+\begin{table}[h]
+\begin{tabular}{lr}
+dogs & 27\% \\
+cats & 48\% \\
+fish & 10\% \\
+bird & 8\% \\
+small mammal & 4\% \\
+reptile & 3\%
+\end{tabular}
+\end{table}
+
 For the first question, we group cats on one side and everything else on the other, to form a 48/52 split.
@@
 If all permutations are equally likely, then $\log W (N_i)$ represents the number of bits needed to identify one of the sequences.
-Assuming we use base $b$ for the logarithm, we take $N$ to be large and, using Stirling's approximation $\ln N!
-\approx N \ln N - N$, we find:
-\begin{equation}
-\begin{aligned}
-\log W(N_i) &= \log \frac{N!}{\prod_{i \in I} N_i!} \\
-&= \log N! - \sum_{i \in I} \log N_i! \\
-&\approx \frac{1}{\ln b} \left(N \ln N - N - \sum_{i \in I} N_i \ln N_i + \sum_{i \in I} N_i \right) \\
-&= \frac{1}{\ln b} \left(\sum_{i \in I} N_i \ln N - \sum_{i \in I} N_i \ln N_i \right) \\
-&= -\sum_{i \in I} N_i \left(\log N_i - \log N \right) \\
-&= - N \sum_{i \in I} \frac{N_i}{N} \log \frac{N_i}{N} \\
-&= - N \sum_{i \in I} p_i \log p_i \\
-&= N H(p_i)
-\end{aligned}
-\end{equation}
 
+When $N$ is large, we can use Stirling's approximation $\ln N!
+\approx N \ln N - N$, and find $\log W(N_i) \approx - N \sum_{i \in I} \frac{N_i}{N} \log \frac{N_i}{N} = N H(p_i)$.
+The logarithm of the permutations is $N$ times the Shannon variability.
 The result should not be surprising: it simply tells us that encoding a sequence of $N$ elements is the same as encoding $N$ elements one at a time.
@@
 The distribution is not what is being encoded.
-The variability is the information needed to go from the distribution, which is given, to an individual element.
+The variability is quantified by the information needed to go from the distribution, which is given, to an individual element.
 It is really the information gap from the population to an element.
@@
 
-This should suggest that the success of Shannon variability and its use in both information theory and physics does not warrant a fundamental change of perspective in what constitutes a physical object, as some physicists have claimed.
-It is true that any physical process can be used to process information, when properly encoded.
+The use of information in physics, then, does not warrant a fundamental change of perspective in what constitutes a physical object, as some physicists have claimed.
+It is true that any physical process can be used to process information, \emph{when properly encoded}.
 It is also true that scientific theories, in the end, are models that can capture only the aspects of nature that can be tested experimentally, the information extracted by the experiment, under suitable circumstances.
-Therefore the claim that information plays an essential role in physical theories has a valid basis.[cite Wheeler, Landauer] But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
+Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value it is absorbed'', "the information thus solicited [by the experiment] make physics and comes in bits" by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations avaialble (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
 The art of experimental science is neither contained in the mathematical description nor in the data collected.
 As information, in the information theoretic sense, requires that context to become intelligible, it cannot play a primary role.
-Therefore the claim that the universe itself is information does not follow.[cite Vedral, Davies]
+Therefore the claim that the universe itself is information does not follow.\footnote{Comments like ``[Information is] the fundamental building block of the Universe'' by Vedral or ``[Information] occupies the ontological basement'' by Davies go in this direction.}
 
@@
 Therefore, instead of having a discrete normalized distribution $\sum_i p_i = 1$, we have a continuous normalized distribution $\int \rho(x)dx=1$.
-While simply changing $- \sum_i p_i \log p_i$ to $- \int \rho(x) \log \rho(x) dx$ works in most cases, it is easy to get confused because the two expressions have significantly different properties.
+While simply changing $- \sum_i p_i \log p_i$ to $- \int \rho(x) \log \rho(x) dx$ works in most cases, leaving it at that misses a crucial point: the two expressions have significantly different properties.
 
@@
 
-This brings up a second problem: what happens if we change unit? If we change the reference, we expect that the entropy will change accordingly. In fact:
+This brings up a second problem: what happens if we change unit? If we change the reference, we expect that the entropy will change accordingly. In fact we find
 \begin{equation}
 \begin{aligned}
-H(\rho(y)) &= - \int \rho(y) \log \rho(y) dy \\
-&= - \int \rho(x) \left|\frac{\partial y}{\partial x}\right| \log \left(\rho(x)\left|\frac{\partial y}{\partial x}\right| \right) \left|\frac{\partial x}{\partial y}\right| dx \\
-&= - \int \rho(x)  \log \left(\rho(x)\left|\frac{\partial y}{\partial x}\right| \right)  dx \\
-&= - \int \rho(x)  \log \rho(x)  dx - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
-&= H(\rho(x)) - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
+H(\rho(y)) &= H(\rho(x)) - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx 
 \end{aligned}
@@
 
-In physics, there is one thing that makes the Shannon variability work: phase-space.
-What happens in phase space is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components.
+In physics, what makes the Shannon variability work is phase-space.
+What happens is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components.
 Therefore the areas $d\hat{q}^i d\hat{p}_i = dq^j dp_j$ remain the same.
@@
 While this is true, their invariance under coordinate transformations is more important as it is what makes them physically meaningful in the first place.
-In fact, how could we have conservation over time if we could not compare it in an observer independent way?
 This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\int \rho \log \rho$.
@@
 
-As we saw in the previous sections, Shannon variability is a very general concept that applies to any type of distribution, no matter what the problem domain.
-It does not require us to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamics and statistical mechanics.
-Yet it can, and is, applied successfully in statistical mechanics even in systems far from equilibrium.[cite] How does this work?
-
-As statistical mechanics aims to describe the collective behavior of a large number of physical systems (i.e.
-particles), the Shannon variability can be applied to their distribution and will correspond to the entropy of the system.
+Statistical mechanics aims to describe the collective behavior of a large number of physical systems (i.e.
+particles), therefore asking what variability is expressed by such a collection is a well posed question.
 Yet, we have to understand that there are two distinct ways to apply the concept simply because there are two types of distributions one can be interested in.
@@
 This will lead to the Gibbs or the Von Neumann entropy depending on whether the system is classical or quantum.
-Both these types of distributions are used and are therefore of interest, so we will look at both and clarify the potential sources of confusion.
+Both these types of distributions are used and are therefore of interest, so we will look at both.
 
@@
 As we said before, we treat continuous variables by comparing to a finite unit.
-We can pick a unit of phase space small enough such that the density $\rho$ is constant over cells of that size.
+We can pick a unit of phase space small enough such that the density $\rho$ can be considered constant over cells of that size.
 We express $\rho$ as the number of particles within the chosen unit and divide phase space into cells.
@@
 
-This setting, distributions over single particle phase space, is the one used to derive the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the distribution of particles for an ideal gas.
-It provides a clear picture for what happens in simple cases: the equilibrium distribution is the one where the variability of the states of the particles is maximal.
-That is, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles.
-The information needed to identify a particle within the distribution at a given precision will increase during an irreversible process.
-The equilibrium is a statistical equilibrium, particles are moving around, but for one particle that moves in one direction, there is another one that moves in the opposite and the overall distribution remains the same.\footnote{This corresponds to the insight developed by Boltzmann that led to the creation of statistical mechanics.}
+Distributions over single-particle phase space is the setting used when deriving the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the distribution of particles for an ideal gas.
+As it evolves towards equilibrium, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles, increasing the variability until is is maximized.
+The equilibrium is a statistical equilibrium, particles are moving around, but for one particle that moves in one direction, there is another one that moves in the opposite and the overall distribution remains the same.\footnote{This corresponds to the original insights developed by Boltzmann.}
 
-This approach, though, will only work in the limit of a large number of indistinguishable particles that are independently distributed.
-If $N$ is not large, there may not be another particle that moves in the opposite direction.
-If there are correlations, the motions may not cancel out.
-More precisely, suppose we have a probability distribution $\hat{\rho}$ for $N$ particles.
+This approach, though, will only work in the limit of a fixed large number of indistinguishable particles that are independently distributed.
+More precisely, suppose we have a joint probability distribution $\hat{\rho}$ for $N$ particles.
 If they are independently distributed, then the joint probability $\hat{\rho}=\prod_{i}\rho_i$ is the product of the distribution for each particle.
@@
 If $N$ is large, the number of particles in one region $N_\Sigma$ is very close to the expectation value $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$.
-If these assumptions are not met, we cannot break the joint distribution into single particle ones, the fluctuations may become relevant thus requiring a more general account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.} Moreover, a change of the number of particles or particles of different species cannot be captured by a single normalized distribution over single particle states.
+If these assumptions are not met, we cannot break the joint distribution into single particle ones, particle number and type may change, the fluctuations may become relevant thus requiring a more general account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.}
 
 The more general setting, then, is the following.
-The macrostate is a probability distribution over all possible complete descriptions of the system, the microstates.
+The macroscopic state, or macrostate, is a probability distribution over all possible complete descriptions of the system, or microstates.
 That is, we have a distribution over the $6N$ dimensional phase space of $N$ particles, sometimes referred to as $\Gamma$-space, where each point represents the position and momentum of $N$ particles.
@@
 As we said, this would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon variability.
-It is more appropriate, both mathematically and conceptually, to think of indistinguishable particles as identical peaked distributions, each contributing the same amount of Shannon entropy.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.} The position and momentum corresponds to the center of mass of the particle.
+It is more appropriate, both mathematically and conceptually, to think of particles as identically peaked distributions, each characterized by the same amount of Shannon variability.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.} The position and momentum corresponds to the center of mass of the particle.
 
-We note that some authors choose to interpret the probability distribution not as representing the frequency over a large amount of independent trials, but as the knowledge one has about the system.
-This would make the entropy an observer dependent quantity, instead of an objective notion.\footnote{Jaynes called it the anthropomorphic nature of entropy.} The issue is that, experimentally, the fluctuations we observe are not observer dependent.
-We believe the confusion comes from the, correct, realization that the system by itself is not enough to determine the thermodynamic variables that are needed to reach the equilibrium.
-In fact, Jaynes points out:
+We note that some authors choose to interpret the probability distribution not as coming from repeated independent trials, but as the knowledge one has about the system.
+This would make the entropy a subjective notion\footnote{Jaynes called it the anthropomorphic nature of entropy.}, but the associated fluctuations we experimentally observe are indeed objective.
+We believe the confusion comes from the, correct, realization that the same system, under different conditions, will be described with a different set of thermodynamic variables.\footnote{Jaynes points out:
 \begin{quote}
@@
 It is clearly  meaningless to ask, ``What is the entropy of the crystal?''  unless we first specify the set of parameters which define its thermodynamic state.
-\end{quote}
-This is a particular case of what we stated in the previous sections: a distribution, and therefore its variability, is contingent upon some arbitrary choices.
-But, again, once those choices are made, the distribution is objective.
+\end{quote}}
 The choice of the system, in statistical mechanics and thermodynamics, is enough to determine the state space for the microstates, but not enough to determine the set of macrostates that correspond to equilibria.
 We have to specify the process and the constraints that that process puts on the system.
-That is, under a different choice of process and constraints the microstates will fluctuate in different ways since we have changed the dynamics of the system.
+Under a different choice of process and constraints the microstates will fluctuate in different ways since we have changed the dynamics of the system.
+This is what we stated at the beginning: a distribution, and therefore its variability, is always contingent upon some arbitrary choices.
 The question, then, is whether the system plus the process plus the set of constraints are enough to determine the entropy, or if we also need to mix in human knowledge.
@@
 The different entropies in statistical mechanics, then, all have a tight link to the Shannon variability.
-The Boltzmann entropy corresponds to the variability of the state of a particle within a given microstate, provided that there are a large number of independent and indistinguishable particles.
+The Boltzmann entropy corresponds to the variability of the state of a particle within a given microstate, provided that there are a large fixed number of independently distributed and indistinguishable particles.
 The Gibbs entropy corresponds to the variability of a classical microstate as constrained by the macrostate.
@@
 The characterization we gave to the Shannon formula, then, is readily applicable to statistical mechanics in a natural way.
-We find it more precise than the common characterizations, such as disorder or lack of knowledge, and, once internalized, it leads to less confusion.
 
 \section{Conclusion}
+
+In this paper we have seen that:
+\begin{itemize}
+	\item the Shannon entropy measures the variability of the elements within a distribution, giving it a crisp intuitive meaning that is a general and applicable to all branches of science
+	\item the expression is not arbitrary, as it is the only linear indicator for such concept
+	\item it measures the variability by quantifying the number of questions one must ask to identify an element within the distribution, which corresponds to the number of bits needed to transmit or store that information
+	\item when properly applied to statistical mechanics, the Shannon entropy leads to the Boltzmann, Gibbs and Von Neumann entropies.
+\end{itemize}
+The characterization we gave to the Shannon formula, then, is more precise than the common characterizations, such as disorder, information or lack of knowledge, and we find it leads to less confusion.
+It clarify that the Shannon variability is an independence concept from the entropy of thermodynamics and statistical mechanics, and a link can be recovered only if properly applied.
+We find that this approach, once internalized, gives greater intuitive insight which also maps more readily with the mathematical details.
 

### 2019-11-19 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
Added Christine comments

@@
 In this paper we will show that the Shannon entropy can be fully understood as measuring the variability of the elements within a distribution: it characterizes how much variation can be found within a collection of objects.
-We will see that it is the only indicator that is continuous and linear, that quantifies the number of yes/no questions (i.e.
-bits) that are needed to identify an element within the distribution, and we will see how applying this concept to statistical mechanics in different ways lead to the Boltzman, Gibbs and Von Neumann entropies.
+We will see that it is the only indicator that is continuous and linear, that quantifies the number of yes/no questions (i.e.~bits) that are needed to identify an element within the distribution, and we will see how applying this concept to statistical mechanics in different ways lead to the Boltzman, Gibbs and von Neumann entropies.
 \end{abstract}
@@
 It measures the number of questions one must ask to identify an element of the distribution, linking its use to information theory.
-We will study how the expression works over continuous variables, and show how phase-space is special as it leaves the Shannon variability invariant under change of coordinates.
-We will turn to statistical mechanics and apply Shannon variability in different ways to recover connections to the Boltzmann, Gibbs and Von Neumann entropies.
+We will study how the expression works over continuous variables, and show how phase space is special as it leaves the Shannon variability invariant under change of coordinates.
+We will turn to statistical mechanics and apply Shannon variability in different ways to recover connections to the Boltzmann, Gibbs and von Neumann entropies.
 
@@
 This also means that, whatever properties those elements will have, they will represent a matter of fact about that set.
-In turn, this will make some choices arguably more appropriate or interesting than others.
+In turn, this will make some choices arguably more appropriate or insightful than others.
 For example, in taxonomy the set of all animals that have feathers may be more appropriate than the set of all animals that can fly: the first includes only birds while the second includes most birds, a few mammals and a lot of insects.
@@
 
-Having decided what elements to study and the level of description we are interested in, we bin them; that is, we group them based on that description, disregarding the identity of the particular element.
+Having decided what elements to study and the level of description in which we are interested, we bin them; that is, we group them based on that description, disregarding the identity of the particular element.
 What we will be interested in is only the relative frequency $p_i = N_i / N = \left| \{e_\alpha \in E | q(e_\alpha) = q_i \} \right| / N$ of the elements within each bin.
@@
 
-As we noted before, the level at which we describe the elements is not absolute and can change.
+As noted before, the level at which we describe the elements is not absolute and can change.
 For example, we may choose to group the animals in the Gal\'{a}pagos islands first by class (e.g.
@@
 \begin{enumerate}
-\item There is 50\% chance you won one million dollars in the lottery and 50\% chance you didn't.
+\item There is 50\% chance you won one million dollars in the lottery and 50\% chance you won nothing.
 \item There is 50\% chance you won one million dollars in the lottery and 50\% chance you won half a million dollars.
@@
 Not all questions will be able to extract the same amount of information.
-Some questions, like ``is it an animal?" would be redundant.
-Others, like ``is it an American Flamingo (Phoenicopterus ruber)?", would give us a lot of information in the positive case but little in the negative case.
+Some questions, like, ``Is it an animal?" would be redundant.
+Others, like, ``Is it an American Flamingo (Phoenicopterus ruber)?", would give us a lot of information in the positive case but little in the negative case.
 However, there has to be a minimum number of questions that must be asked to get to the answer.
@@
 We can also pick a non-integer base, like the natural base for logarithms, and we will have nats.
-This is why the Shannon variability is fundamental in information theory, because it quantifies how much information is needed to transfer a value picked from a distribution.
+This is why the Shannon variability is fundamental in information theory, because it quantifies \emph{how much information is needed to transfer a value picked from a distribution}.
 
@@
 For the first question, we group cats on one side and everything else on the other, to form a 48/52 split.
-So we can ask ``is it a cat?".
+So we can ask, ``Is it a cat?".
 If the answer is yes, we finished.
@@
 We can group dogs on one side and everything else on the other to form a 27/25 split.
-So we can ask `` is it a dog?".
+So we can ask, ``Is it a dog?".
 If the answer is yes, we are done.
@@
 We can group fish with reptile and bird with small mammal to form a 13/14 split.
-So we can ask ``is it a fish or a reptile?" If the answer is yes, the followup question would be ``is it a fish?" If the answer is no, the followup question would be ``is it a bird?" With this scheme, the encoding, where 1 represents 'yes' and 0 represents 'no' to the each of the questions asked, is as follows:
+So we can ask, ``Is it a fish or a reptile?".
+If the answer is yes, the followup question would be ``Is it a fish?".
+If the answer is no, the followup question would be ``Is it a bird?".
+With this scheme, the encoding, where 1 represents 'yes' and 0 represents 'no' to the each of the questions asked, is as follows:
 \begin{itemize}
@@
 These are taken from $I$ different cases: dogs, cats, fish and so on.
-Suppose $N_i$ are the of elements of each type, which means $\sum_{i \in I} N_i = N$.
+Suppose $N_i$ are the number of elements of each type, which means $\sum_{i \in I} N_i = N$.
 Then, given a particular instance of $N_i$, we have $W(N_i) = \frac{N!}{\prod_{i \in I} N_i!}$ possible ways to realize that sequence, which corresponds to all possible permutations.
@@
 It is really the information gap from the population to an element.
-That is why some people say the entropy is ``lack of information", which is justified because, in a way, it is the information about the elements that the distribution cannot provide.
-But, again, this is just missing the point of what is being described.
-If one is not interested in identifying elements there is no ``missing information".
+That is why some people say the entropy is ``lack of information," which is justified because, in a way, it is the information about the elements that the distribution cannot provide.
+But, again, this is deceptive: if one is not interested in identifying elements there is no ``missing information."
 
@@
 It is also true that scientific theories, in the end, are models that can capture only the aspects of nature that can be tested experimentally, the information extracted by the experiment, under suitable circumstances.
-Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value it is absorbed'', "the information thus solicited [by the experiment] make physics and comes in bits" by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations avaialble (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
-The art of experimental science is neither contained in the mathematical description nor in the data collected.
+Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value it is absorbed'', ``the information thus solicited [by the experiment] make physics and comes in bits'' by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations avaialble (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
+The art of experimental science is contained neither in the mathematical description nor in the data collected.
 As information, in the information theoretic sense, requires that context to become intelligible, it cannot play a primary role.
-Therefore the claim that the universe itself is information does not follow.\footnote{Comments like ``[Information is] the fundamental building block of the Universe'' by Vedral or ``[Information] occupies the ontological basement'' by Davies go in this direction.}
+Therefore the claim that the universe itself is information\footnote{Comments like ``[Information is] the fundamental building block of the Universe'' by Vedral or ``[Information] occupies the ontological basement'' by Davies go in this direction.} does not follow.
 
@@
 How does it work in that case?
-Wouldn't we need infinitely many bits to identify a value from a continuous distribution?
+Would we not need infinitely many bits to identify a value from a continuous distribution?
 
-\section{Continuous distributions}\label{sec_cont}
+\section{Continuous Distributions}\label{sec_cont}
 
-A standard quick and dirty way to extend distributions from discrete variables to continuous variables is to substitute weights with densities and the sums with integrals.
+A standard quick and dirty way to extend distributions from discrete variables to continuous variables is to substitute weights with densities and sums with integrals.
 Therefore, instead of having a discrete normalized distribution $\sum_i p_i = 1$, we have a continuous normalized distribution $\int \rho(x)dx=1$.
@@
 This means that infinitesimally narrow distributions, like delta functions, do not work well with the Shannon variability as they would give minus infinite entropy.
-If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which have a very small but still finite support (i.e.
-region where the function is non-zero).
-We will assume all our continuous distributions are like that.
+If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which have a very small but still finite support (i.e.~region where the function is non-zero), which is what we will assume.
 
-This brings up a second problem: what happens if we change unit? If we change the reference, we expect that the entropy will change accordingly. In fact we find
+This brings up a second problem: what happens if we change units? If we change the reference, we expect that the entropy will change accordingly. In fact we find
 \begin{equation}
@@
 
-In physics, what makes the Shannon variability work is phase-space.
+In physics, what makes the Shannon variability work is phase space.
 What happens is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components.
 Therefore the areas $d\hat{q}^i d\hat{p}_i = dq^j dp_j$ remain the same.
-The Jacobian determiants are unitary and the Shannon variability is invariant.
+The Jacobian determinants are unitary and the Shannon variability is invariant.
 
-The use of phase-space in statistical mechanics is of paramount importance: over this space both the density $\rho(q^i, p
+The use of phase space in statistical mechanics is of paramount importance: over this space both the density $\rho(q^i, p
 _i)$ and its variability $H(\rho(q^i, p
@@
 
-The importance of phase space is often attributed to Liouville's theorem, that states that areas are conserved under Hamiltonian evolution.
+The importance of phase space is often attributed to Liouville's theorem, which states that areas in phase space are conserved under Hamiltonian evolution.
 While this is true, their invariance under coordinate transformations is more important as it is what makes them physically meaningful in the first place.
-This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\int \rho \log \rho$.
-If we use distributions over position and velocities or over momentum only, the expression would change for those particular variables in those coordinate systems.
+This means that we always have to consider distributions over the full phase space with the appropriate conjugate coordinates if we want to use $-\int \rho \log \rho$.
+If we use distributions over position and velocities or over momentum only, the expression will change for those particular variables in those coordinate systems.
 This can be, again, a source of confusion.\cite{Dunkel}
 
-We conclude noting that in quantum mechanics the Shannon variability, in the form of the Von Neumann entropy, is coordinate invariant as well.
+We conclude noting that in quantum mechanics the Shannon variability, in the form of the von Neumann entropy, is coordinate invariant as well.
 That is, given a density matrix $\rho$ then $S_{VN} = - \textrm{tr}(\rho \log \rho)$ is independent of the basis in which it is calculated.
 
-Now that we have seen how the Shannon variability works in the continuous case and the importance of phase space, we are ready to see how it can be used in statistical mechanics and how it relates to the Boltzman, Gibbs and Von Neumann entropies.
+Now that we have seen how the Shannon variability works in the continuous case and the importance of phase space, we are ready to see how it can be used in statistical mechanics and how it relates to the Boltzmann, Gibbs and von Neumann entropies.
 
@@
 
-Statistical mechanics aims to describe the collective behavior of a large number of physical systems (i.e.
-particles), therefore asking what variability is expressed by such a collection is a well posed question.
-Yet, we have to understand that there are two distinct ways to apply the concept simply because there are two types of distributions one can be interested in.
+Statistical mechanics aims to describe the collective behavior of a large number of physical systems (i.e.~particles), therefore asking what variability is expressed by such a collection is a well posed question.
+Yet, we have to understand that there are two distinct ways to apply the concept simply because there are two types of distributions in which one can be interested.
 We may consider the state of the whole system at a given time, and be interested in the distribution of the particles over all possible particle states.
-This will lead to the Boltzman entropy.
-We may consider a statistical ensemble, which is a large collection of independent copies of the system found in different states, and consider the distribution of these different instances over the states of the whole system.
-This will lead to the Gibbs or the Von Neumann entropy depending on whether the system is classical or quantum.
+This will lead to the Boltzmann entropy.
+We may instead consider a statistical ensemble, which is a large collection of independent copies of the system found in different states, and consider the distribution of these different instances over the states of the whole system.
+This will lead to the Gibbs or the von Neumann entropy depending on whether the system is classical or quantum.
 Both these types of distributions are used and are therefore of interest, so we will look at both.
@@
 Suppose we have a large number $N$ of particles taken from a normalized distribution $\rho(q^i, p_i)$.
-The space is the six dimensional phase-space for a single particle, sometimes called $\mu$-space.
+The space is the six-dimensional phase space for a single particle, sometimes called $\mu$-space.
 As we assume $N$ large, we can think of $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$ as the number of particles that are within a region $\Sigma$ of phase space.
@@
 As it evolves towards equilibrium, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles, increasing the variability until is is maximized.
-The equilibrium is a statistical equilibrium, particles are moving around, but for one particle that moves in one direction, there is another one that moves in the opposite and the overall distribution remains the same.\footnote{This corresponds to the original insights developed by Boltzmann.}
+The equilibrium is a statistical equilibrium, particles are moving around, but for any particle that moves in one direction, there is another one that moves in the opposite and the overall distribution remains the same.\footnote{This corresponds to the original insights developed by Boltzmann.}
 
@@
 If $N$ is large, the number of particles in one region $N_\Sigma$ is very close to the expectation value $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$.
-If these assumptions are not met, we cannot break the joint distribution into single particle ones, particle number and type may change, the fluctuations may become relevant thus requiring a more general account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.}
+If these assumptions are not met, we cannot break the joint distribution into single-particle ones, particle number and type may change, the fluctuations may become relevant thus requiring a more general account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.}
 
@@
 The macroscopic state, or macrostate, is a probability distribution over all possible complete descriptions of the system, or microstates.
-That is, we have a distribution over the $6N$ dimensional phase space of $N$ particles, sometimes referred to as $\Gamma$-space, where each point represents the position and momentum of $N$ particles.
+That is, we have a distribution over the $6N$-dimensional phase space of $N$ particles, sometimes referred to as $\Gamma$-space, where each point represents the position and momentum of $N$ particles.
 The Gibbs entropy is $S_G = -k_B \int \rho \log \rho \, dq^idp_i$, which corresponds to the Shannon variability of the microstate distribution for a given classical macrostate.
@@
 
-There are a couple of problems in this picture that we have to be aware of.
+There are a couple of issues in this picture.
 The first is that $\Gamma$-space automatically assumes that all particles are distinguishable.
 This leads to the widely known problem of overcounting which needs to be addressed in the standard way.[cite] The second problem is that, though the state of each particle is given by position and momentum, we should not think of them as literally pointlike.
-As we said, this would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon variability.
-It is more appropriate, both mathematically and conceptually, to think of particles as identically peaked distributions, each characterized by the same amount of Shannon variability.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.} The position and momentum corresponds to the center of mass of the particle.
+As we said, this would correspond to delta Dirac distributions over phase space which have minus infinite Shannon variability.
+It is more appropriate, both mathematically and conceptually, to think of particles as identically peaked distributions, each characterized by the same amount of Shannon variability.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.} The position and momentum correspond to the center of mass of the particle.
 
@@
 The case of a quantum system is formally similar to the classical one.
-Instead of a distribution over the $N$ particle phase space, we have a distribution over the Hilbert space for the quantum system which is represented by a density matrix operator $\rho$.
-The Von Neumann entropy is given by $S_{VN} = - \textrm{tr}(\rho \log \rho)$.
-This, expanded in a basis, becomes $S_{VN} = - \sum_i \rho_i \log \rho_i$ or $S_{VN} = - \int \rho(x) \log \rho(x)dx$ depending on whether the spectrum is discrete or continuous.
-The Von Neumann entropy corresponds to the Shannon variability of the microstate distribution for a given quantum macrostate.
+Instead of a distribution over the $N$-particle phase space, we have a distribution over the Hilbert space for the quantum system which is represented by a density matrix operator $\rho$.
+The von Neumann entropy is given by $S_{V\!N} = - \textrm{tr}(\rho \log \rho)$.
+This, expanded in a basis, becomes $S_{V\!N} = - \sum_i \rho_i \log \rho_i$ or $S_{V\!N} = - \int \rho(x) \log \rho(x)dx$ depending on whether the spectrum is discrete or continuous.
+The von Neumann entropy corresponds to the Shannon variability of the microstate distribution for a given quantum macrostate.
 
@@
 The Gibbs entropy corresponds to the variability of a classical microstate as constrained by the macrostate.
-The Von Neumann entropy is similar but corresponds to the quantum case.
+The von Neumann entropy is similar but corresponds to the quantum case.
 The characterization we gave to the Shannon formula, then, is readily applicable to statistical mechanics in a natural way.
@@
 	\item it measures the variability by quantifying the number of questions one must ask to identify an element within the distribution, which corresponds to the number of bits needed to transmit or store that information
-	\item when properly applied to statistical mechanics, the Shannon entropy leads to the Boltzmann, Gibbs and Von Neumann entropies.
+	\item when properly applied to statistical mechanics, the variability leads to the Boltzmann, Gibbs and von Neumann entropies.
 \end{itemize}
-The characterization we gave to the Shannon formula, then, is more precise than the common characterizations, such as disorder, information or lack of knowledge, and we find it leads to less confusion.
+The characterization we gave to the Shannon formula, then, is more precise than the common characterizations, such as disorder, information or lack of knowledge, and it should lead to less confusion.
 It clarify that the Shannon variability is an independence concept from the entropy of thermodynamics and statistical mechanics, and a link can be recovered only if properly applied.

### 2019-11-19 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
Update AJP-ShannonEntropyIsVariability.tex

@@
 	The Shannon entropy, one of the cornerstones of information theory, is widely used in physics, particularly in statistical mechanics.
-Yet its characterization and connection to physics remains vague, leaving ample room for misconceptions and misunderstanding.
-In this paper we will show that the Shannon entropy can be fully understood as measuring the variability of the elements within a distribution: it characterizes how much variation can be found within a collection of objects.
-We will see that it is the only indicator that is continuous and linear, that quantifies the number of yes/no questions (i.e.~bits) that are needed to identify an element within the distribution, and we will see how applying this concept to statistical mechanics in different ways lead to the Boltzman, Gibbs and von Neumann entropies.
+Yet its characterization and connection to physics remain vague, leaving ample room for misconceptions and misunderstanding.
+We will show that the Shannon entropy can be fully understood as measuring the variability of the elements within a given distribution: it characterizes how much variation can be found within a collection of objects.
+We will see that it is the only indicator that is continuous and linear, that it quantifies the number of yes/no questions (i.e.~bits) that are needed to identify an element within the distribution, and we will see how applying this concept to statistical mechanics in different ways leads to the Boltzmann, Gibbs and von Neumann entropies.
 \end{abstract}
@@
 The Shannon entropy $-\sum p_i \log p_i$ is well established in such disparate fields as computer science[cite], physics[cite] and ecology[cite].
-It is most often introduced in physics through statistical mechanics, and associated to vague notions, such as uncertainty[cite], knowledge[cite], lack of knowledge[cite] and disorder[cite].
-These prevent a proper understanding of the concept and often lead to confusion and misunderstandings.
+It is most often introduced in physics through statistical mechanics, and associated with vague notions, such as uncertainty[cite], knowledge[cite], lack of knowledge[cite] and disorder[cite].
+These prevent a proper understanding of the concept and often lead to confusion.
 
 This paper aims to give a crisp characterization of the Shannon entropy that is intuitive and precise.
-The general idea is that it measures the variability of the elements within a distribution, which is a general concept applicable to many branches of science and it is independent of the notion of entropy of thermodynamics and statistical mechanics.
-Therefore we will use the term \emph{``Shannon variability''}, leaving ``entropy'' to the physical concept.
+The general idea is that it measures the variability of the elements within a distribution, which is a general concept applicable to many branches of science, and it is independent of the notion of entropy of thermodynamics and statistical mechanics.
+Therefore we will use the term \emph{``Shannon variability,''} leaving ``entropy'' to the physical concept.
 We will show that the expression $- \sum_i p_i \log p_i$ is the only linear indicator of variability.
@@
 Therefore we assume little knowledge of the subject from the reader and include all details that we feel are important to properly understand the subject without confusion, including two standard derivations for the formula and examples that are common in computer science but rarely used in physics.
-We will briefly touch over a few of the common misconceptions.
+We will briefly touch on a few of the common misconceptions.
 
@@
 We have a set of elements $E = \{e_\alpha\}_{\alpha=1}^N$.
-In accordance with the comments made in our introduction, they can for example represent the galaxies in the Laniakea Supercluster, the animals in the Gal\'{a}pagos islands, the molecules in a box of gas given a particular macrostate, the different outcomes of spin measurement for a given quantum state, or the words in the complete works of Shakespeare.
+In accordance with the comments made in the introduction, they can for example represent the galaxies in the Laniakea Supercluster, the animals in the Gal\'{a}pagos islands, the molecules in a box of gas given a particular macrostate, the different outcomes of spin measurement for a given quantum state, or the words in the complete works of Shakespeare.
 The choice of the particular set is driven by the interests and practices of scientists and their fields of study and therefore it is, in this sense, discretionary.
@@
 Once we arbitrarily decided we will study the galaxies in the Laniakea Supercluster, it is a matter of fact that the Milky Way will be included.
-This also means that, whatever properties those elements will have, they will represent a matter of fact about that set.
+This also means that, whatever properties those elements have, they will represent a matter of fact about that set.
 In turn, this will make some choices arguably more appropriate or insightful than others.
@@
 
-Once we have constructed our distribution, we will want to construct an indicator $H$ that quantifies how much variability the elements exhibit within the distribution.
+Once we have constructed our distribution, we want to construct an indicator $H$ that quantifies how much variability the elements exhibit within the distribution.
 That is, we want to quantify the degree of diversity that the values can have within the distribution.
@@
 types of galaxies, words), such statistical quantities are not well defined.
-Second, relabeling the values (e.g.
-switching names, switching units, non-linearly changing coordinates) does not have an impact on variability of the elements, while statistical quantities will in general be affected.
+Second, relabeling the values (e.g.~switching names, switching units, non-linearly changing coordinates) does not have an impact on variability of the elements, while statistical quantities will in general be affected.
 This tells us that our indicator cannot depend on $Q$ itself, but only on the weights $p_i$.
@@
 Though this is a source of additional confusion, for the purpose of defining a measure of variability it does not change things conceptually.
-We will discuss the problem in section \ref{sec_cont}.}
+We discuss the issue in section \ref{sec_cont}.}
 
@@
 As noted before, the level at which we describe the elements is not absolute and can change.
-For example, we may choose to group the animals in the Gal\'{a}pagos islands first by class (e.g.
-mammals, birds, reptiles) and then later refine the mammals by species.
+For example, we may choose to group the animals in the Gal\'{a}pagos islands first by class (e.g.~mammals, birds, reptiles) and then later refine the mammals by species.
 In this case, we would like $H$ to combine linearly with respect to the weights.
@@
 \end{enumerate}
-These are the same requirements Shannon put forth for his expression\cite{Shannon}, from which he showed to be the only possible choice: $H(p_i) = - \sum p_i \log(p_i)$.
+These are the same requirements Shannon put forth for his expression\cite{Shannon}, from which he showed that the only possible choice is $H(p_i) = - \sum p_i \log(p_i)$.
 This expression quantifies the variability of the elements within the distribution, the variety of values one finds.
@@
 \end{enumerate}
-The distribution in both cases is the same, two bins 50\% chance each, and so is the Shannon variability. Yet, you know more in the second case: you know you have won at least half a million dollars.
+The distribution in both cases is the same, two bins 50\% chance each, and so is the Shannon variability. Yet, you know more in the second case: you know you won at least half a million dollars.
 
@@
 If we look at the galaxies in our universe, what variability do they exhibit in terms of their types?
-If we look at the animals in the Gal\'{a}pagos island, what variability is expressed in terms of their species?
+If we look at the animals in the Gal\'{a}pagos islands, what variability is expressed in terms of their species?
 If we look at the molecules in a given macrostate, what variability do they express in terms of their microstates?
@@
 Suppose we fix a distribution, say the animals in the Galap\'{a}gos islands binned by their respective species.
-Suppose I pick a specific animal from the set and you want to know its species.
+Suppose we pick a specific animal from the set and you want to know its species.
 Suppose the only way for you to get that information is to ask a series of questions with only two possible answers, yes or no.
@@
 We can also pick a non-integer base, like the natural base for logarithms, and we will have nats.
-This is why the Shannon variability is fundamental in information theory, because it quantifies \emph{how much information is needed to transfer a value picked from a distribution}.
+This is why the Shannon variability is fundamental in information theory, because it quantifies \emph{how much information is needed to transfer a value picked from a known distribution}.
 
@@
 fish & 10\% \\
-bird & 8\% \\
-small mammal & 4\% \\
-reptile & 3\%
+birds & 8\% \\
+small mammals & 4\% \\
+reptiles & 3\%
 \end{tabular}
@@
 
+\noindent
 For the first question, we group cats on one side and everything else on the other, to form a 48/52 split.
@@
 If not, we continue.
-We can group fish with reptile and bird with small mammal to form a 13/14 split.
+We can group fish with reptiles and birds with small mammals to form a 13/14 split.
 So we can ask, ``Is it a fish or a reptile?".
@@
 \item fish 10\% - 4 questions (i.e.~4 bits) - answers: [no, no, yes, yes] (i.e.~encoding 0011)
-\item bird 8\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, yes] (i.e.~encoding 0001)
-\item small mammal 4\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, no] (i.e.~encoding 0000)
-\item reptile 3\% - 4 questions (i.e.~4 bits) - answers: [no, no, yes, no] (i.e.~encoding 0010)
+\item birds 8\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, yes] (i.e.~encoding 0001)
+\item small mammals 4\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, no] (i.e.~encoding 0000)
+\item reptiles 3\% - 4 questions (i.e.~4 bits) - answers: [no, no, yes, no] (i.e.~encoding 0010)
 \end{itemize}
@@
 It is also true that scientific theories, in the end, are models that can capture only the aspects of nature that can be tested experimentally, the information extracted by the experiment, under suitable circumstances.
-Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value it is absorbed'', ``the information thus solicited [by the experiment] make physics and comes in bits'' by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations avaialble (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
+Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value it is absorbed'', ``the information thus solicited [by the experiment] make physics and comes in bits'' by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations avaialble (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context in order to be understood: we need to know what the subject of our experiment is, how to prepare it and how to collect the data.
 The art of experimental science is contained neither in the mathematical description nor in the data collected.
@@
 This will lead to the Gibbs or the von Neumann entropy depending on whether the system is classical or quantum.
-Both these types of distributions are used and are therefore of interest, so we will look at both.
+Both these types of distributions are used and are therefore of interest, so we will examine both.
 
@@
 
-Distributions over single-particle phase space is the setting used when deriving the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the distribution of particles for an ideal gas.
-As it evolves towards equilibrium, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles, increasing the variability until is is maximized.
+Distributions over single-particle phase space constitute the setting used when deriving the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the distribution of particles for an ideal gas.
+As it evolves towards equilibrium, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles, increasing the variability until it is maximized.
 The equilibrium is a statistical equilibrium, particles are moving around, but for any particle that moves in one direction, there is another one that moves in the opposite and the overall distribution remains the same.\footnote{This corresponds to the original insights developed by Boltzmann.}
@@
 \begin{itemize}
-	\item the Shannon entropy measures the variability of the elements within a distribution, giving it a crisp intuitive meaning that is a general and applicable to all branches of science
+	\item the Shannon entropy measures the variability of the elements within a given distribution, giving it a crisp intuitive meaning that is a general and applicable to all branches of science
 	\item the expression is not arbitrary, as it is the only linear indicator for such concept
-	\item it measures the variability by quantifying the number of questions one must ask to identify an element within the distribution, which corresponds to the number of bits needed to transmit or store that information
+	\item it measures the variability by quantifying the number of yes/no questions one must ask to identify an element within the distribution, which corresponds to the number of bits needed to transmit or store that information
 	\item when properly applied to statistical mechanics, the variability leads to the Boltzmann, Gibbs and von Neumann entropies.
@@
 The characterization we gave to the Shannon formula, then, is more precise than the common characterizations, such as disorder, information or lack of knowledge, and it should lead to less confusion.
-It clarify that the Shannon variability is an independence concept from the entropy of thermodynamics and statistical mechanics, and a link can be recovered only if properly applied.
-We find that this approach, once internalized, gives greater intuitive insight which also maps more readily with the mathematical details.
+It clarifies that the Shannon variability is an independent concept from the entropy of thermodynamics and statistical mechanics, and a link can be recovered only if properly applied.
+We find that this approach, once internalized, gives greater intuitive insight and also maps more readily to the mathematical details.
 

### 2019-11-20 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
Fixing citations

@@
 It is also true that scientific theories, in the end, are models that can capture only the aspects of nature that can be tested experimentally, the information extracted by the experiment, under suitable circumstances.
-Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value it is absorbed'', ``the information thus solicited [by the experiment] make physics and comes in bits'' by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations avaialble (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context in order to be understood: we need to know what the subject of our experiment is, how to prepare it and how to collect the data.
+Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value'', ``the information thus solicited [by the experiment] makes physics and comes in bits'' by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations avaialble (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context in order to be understood: we need to know what the subject of our experiment is, how to prepare it and how to collect the data.
 The art of experimental science is contained neither in the mathematical description nor in the data collected.
 As information, in the information theoretic sense, requires that context to become intelligible, it cannot play a primary role.
-Therefore the claim that the universe itself is information\footnote{Comments like ``[Information is] the fundamental building block of the Universe'' by Vedral or ``[Information] occupies the ontological basement'' by Davies go in this direction.} does not follow.
+Therefore the claim that the universe itself is information\footnote{Comments like ``[Information is] the fundamental building block of the Universe'' by Vedral\cite{Vedral} or ``[Information] occupies the ontological basement'' by Davies\cite{Davies} go in this direction.} does not follow.
 

### 2019-11-20 Articles/Overview/overview.tex
Starting to work on an overview of the project

\documentclass[twocolumn]{article}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{enumitem}

\title{Overview and Status of the Assumptions of Physics}
\author{Gabriele Carcassi, Christine A. Aidala}
\date{November 2019}

\begin{document}

\maketitle

\begin{abstract}
	This article gives an overview of the project ``Assumptions of Physics'' which aims to rederive the known laws from a few physically meaningful starting points. It presents the motivations behind the project, a summary of the main findings and current status of the research.	
\end{abstract}

\section{Introduction}

The aim of this article is to give a broad overview of the project ``Assumptions of Physics'' so that the reader will understand its motivations, its status and a list of the main findings. This is intended to be a good start for anyone interested in collaborating or simply learning more about the project. Each subsection will be fairly self-contained to allow the reader to skim through the document and focus on the parts of interest.

As we plan to update this document, please make sure this is a recent version.

\subsection{Index}
To this paper, not to the work.

- Goals and method
Goals: Understanding, have a better sense of the math, give a background for new theories...
Methods: reverse engineering vs forward engineering, holistic

- General theory
relationship between math and physics
logic of verifiable statements
properties and quantities
accuracy
differentiability

- Fundamental physics
states and processes
determinism/reversibility: isomorphism in the category
entropy

- Physical theories
Classical mechanics
Quantum mechanics
Thermodynamics
Statistical mechanics

\section{Goals and method}

A formal framework that forces us to clarify our assumptions
From those assumptions the mathematical objects are derived
Each mathematical object has a clear physical meaning and no object is unphysical
Gives us concepts and tools that span across different disciplines
Allows to explore what happens when the assumptions fail, possibly leading to new physics ideas

\textbf{A better understanding of physics.} The overall goal of the project is to better understand physics: to understand why the laws of quantum mechanics, or classical mechanics, are what they are; to understand the limit of the validity of the different theories; to understand what the Poisson brackets, or any other mathematical operation, represent physically; to understand which mathematical objects correspond to actual physical entities and which are just artifacts; to understand what new theories are possible.

To us, understanding means to see the necessity of something and how it relates to the rest. For example, once we know that the earth's revolution axis is not aligned with earth's rotation axis, we understand why there must be seasons; and we understand why days are shorter in the winter; and why the sun is brighter and warmer in the summer. The more connections, the fewer starting points, the more understanding.

\subsection{Objectives}

The overall goal of a better understanding of physics can be broken down into the following five objectives.

\textbf{Clarify the assumptions.} Physics is currently a patchwork of different theories (e.g. classical mechanics, quantum mechanics, thermodynamics, ...) that are used for different systems or for the same system in different contexts. When should we use one instead of another? What are the limits of validity of each theory? We want to understand what are the assumptions one has to make on a system such that it can be studied with a particular theory. The only way to make sure our assumptions are necessary and sufficient is to show the theory can be rederived from them. This, in turn, gives us a better idea of what each theory describes, how it fails, and what new directions one could take.

\textbf{Put physics back at the center of the discussion.} Physics used to try and identifies basic ``laws'' or ``principles'' for itself. Current physical theories, instead, just postulate a mathematical structure relegating physics to a mere after-the-fact interpretation. We want to go back to the old approach: we want to start with the physics and derive the math suitable to capture those physical concepts. Ideally, no mathematically construct should be introduced if it is not physically motivated. Why is phase-space formed with conjugate pairs? What does the commutator represent? If these are really representing physical objects, we should have a better understanding than a vague ``you can think of it as...'' or a cryptic mathematical ``it is the left action of a fiber bundle on a...''. Most of our understanding should come from physical intuition and not from the mathematics it is used to represent it.

\textbf{Give science sturdier mathematical grounds.} Physicists are generally trained to think that mathematical details are physically not interesting and are the domain of mathematicians. Even mathematical structures that originally came from physics are now formalized by mathematicians to solve their own needs, and not the physicist's. This, in turns, reinforces the idea that those details are not interesting to the physicists. The reality is that a well-posed physical problem must also be a well-posed mathematical problem, but not the other way around, so leaving the details to the mathematicians means having fundamental structures that are less physical. For example, quantum states are modeled with Hilbert spaces even though we know they contain mathematical objects that are unphysical. If we give a theory, as we said before, by simply stating a mathematical structure, then there is no guarantee that what we give is a fully physical theory. Unphysical elements simply means the physical problem was not specified correctly. If we truly are able to give a precise physical meaning to every mathematical object, then our formal structures will map one-to-one to our physical understanding. This means no unphysical mathematical artifacts and a more precise mathematical treatment of physics.

\textbf{Foster connections between different fields of knowledge.} Knowledge is increasingly specialized and fewer and fewer people are well versed in more than a couple of subjects. Yet, connections between different fields of science, mathematics and engineering are routinely found to be useful for one simple reason: nature is one and does not care about such divisions. A more holistic vision of scientific knowledge, then, is a natural byproduct of our effort. The notions of state, environment, process and equilibrium, for example, are intertwined and are fundamental to most fields of science and engineering. Proper characterization of such basic concepts will give a common language and mathematical tools that span across different disciplines.

\textbf{Provide a solid basis for new theories.} Most of the attention in fundamental physics is focused on the development of new theory or the search for new effects. A better understanding of the current theories, recast in a single broader and more precise framework, facilitates that search. It gives more insight and intuition leading to new ideas. It makes explicit what ingredients went into the theories, so we know what assumption can fail and what principles cannot be changed. For example, we may abandon the assumption that space is measurable at an indefinite precision, leaving the realm of manifold and real numbers; yet we cannot abandon the notion that space is measurable, staying in the realm of second countable $T_0$ topological spaces. As all the physical theories are part of a more general structure, they also provide templates for new theories and prove that the more general structure is sound.

The objectives are cover many areas and are very broad. As such, we are always looking for people with diverse background and interest. We ultimately believe that all these objectives are connected and cannot be pursued independently.

\subsection{Methodology}

As each topic is investigated by the project, it generally goes through three phases. We mention them here as they will give context to the status of the project and to people who may want to contribute.

\textbf{Reverse engineering.} The goal of this phase is to deconstruct an already existing theory or mathematical structure to understand what each piece is supposed to represent physically. One way to conduct such investigation is to ask what would happen if a particular characteristic of the framework would fail. For example, a line of reasoning could be as follows: Hamilton's equations are differential; if they weren't the mapping of densities and areas would not be well defined; maybe Hamiltonian mechanics is not about mapping points but about mapping densities or areas. This usually leads to a set of necessary assumptions about the physical system one is describing, the ones one must take for the particular description to apply.

\textbf{Forward engineering.} The goal of this phase is to construct an argument from the necessary assumptions outlined in the previous phase to rederive the framework at hand. This checks whether the necessary assumptions are also sufficient. For example, if we started simply by assuming we have density distributions mapped in time, this would give us all possible differentiable evolutions, not necessarily Hamiltonian ones. This leads to a set of necessary and sufficient conditions to rederive the theory.

\textbf{Formalization.} The goal of this phase is to take the previous set of arguments and formalize them in pure mathematical language. This forces us to clarify all the starting points and what parts of the arguments are truly formal and what are physical. This can be done only starting from the foundations and therefore takes a considerable longer time. For example, before formalizing Hamiltonian mechanics we have to formalize differentiable manifolds, real number and topological spaces.

\subsection{Organization}

The overall project is organized into the following macro-categories.

\textbf{The general theory.} This formalizes the basic mathematical framework that is the basis for all science. Like in mathematics every algebraic structure is a specialization of the generic structures axiomatized by set theory, all scientific theory are ultimately specialization of the generic structure provided by the general theory. This defines the basic requirements imposed on a theory by logical consistency and experimental verifiability. It defines properties and quantities, accuracy and other concepts of general applicability.

\textbf{Fundamental physics concepts.} This defines the notions that are fundamental to physics specifically, such as states, processes, environment, equilibrium, determinism and so on. The idea is to define these concepts in the most abstract way possible such that common requirements can be identified. Every physical theory is a further refinement of this structure in that it will choose to study a particular set of states over a particular process.

\textbf{Physical theories.} This studies the different assumptions that are needed to recover the known physical theories. The assumptions typically describe what level of description is available for the system and its parts, how does it change in time and whether trajectories are enough to recover it.

We will present the work in the reverse order, from physical theories to general theory, as we find most physicist

\section{Physics theories}

\subsection{Classical mechanics}

Classical Hamiltonian and Lagrangian particle mechanics, including the Hamiltonian for massive particles under potential forces, can be derived from three main assumptions: infinitesimal reducibility, determinism and reversible evolution and kinematic equivalence. We list here the main points of the derivation together with the some of the insights and connections it provides.

\subsubsection{Classical state spaces}

\textbf{Infinitesimal reducibility.} A system is said reducible if giving the state of the whole system is equivalent to giving the state of its parts and vice-versa. For example, given a ball, we can throw it and study the motion of the ball. Alternatively, one can take a red marker, make a red dot on the ball and study the motion of the dot. The system is reducible if studying the system is equivalent to studying the motion of all possible red dots. The system is infinitesimally reducible if we can keep reducing the system to smaller and smaller parts. We call particle an infinitesimal part of the system, the limit of recursive subdivision.

\emph{Composite states are distributions over particle states.} Under infinitesimal reducibility, if $\mathcal{C}$ is the state space of the while system and $\mathcal{S}$ is the state space of the particles, then each state $c \in \mathcal{C}$ for the full system is identified by a distribution $\rho : \mathcal{S} \to \mathbb{R}$ over the states of infinitesimal parts. The function is real to signify that we can associate arbitrarily small amounts to each particle states. The distribution tells us how many particles can be found in each particle state.

\emph{Continuous time implies particle state space is a manifold.} If time is assumed to be continuous, then the state space $\mathcal{S}$ is a manifold. That is, each particle state $s \in U \subseteq \mathcal{S}$ can be identified by a set of continuous quantities $\xi^a : U \to \mathbb{R}$ which we call state variables. It is a consequence of the general theory that, if we wanted to write a trajectory $\xi^a(t)$, then this must be a topologically continuous function. Therefore, conceptually, once we assume time to be a real number, the state space of the particles must be charted by a set of real quantities: the topology that is given to time reasserts itself on all the quantities that will depend on time.

\emph{Densities imply differentiability.} As we can write the state $s(\xi^a)$ as a function of the state variables $\xi^a$, we will also want to write the density $\rho(s(\xi^a)) = \rho(\xi^a)$ as a function of the state variables. This requires the state variables to be differentiable with respect to the distribution and with each other. Recall that if the Jacobian is not well defined during a change of variable $\rho(\hat{\xi}^b) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a}\right| \rho(\xi^a)$ then the density wouldn't be expressible in the new coordinates. Mathematically, this means that the manifold is equipped with a differentiable structure. Note that any time evolution that would map a density to another density will not only have to be continuous, but also differentiable. In short: assuming density distributions as our primary objects, instead of points, justifies why differentiable structures and differentiable time evolution is present in physics.

\emph{Coordinate invariant densities imply phase-space.} As particles states must be defined independently of coordinate systems, they must be invariant under coordinate transformations. That is, $s(\xi^a) = s(\hat{\xi}^b(\xi^a))$. As the density distribution depends only on the state, it should be invariant as well, that is $\rho(s(\xi^a)) = \rho(s(\hat{\xi}^b(\xi^a)))$. We have two seemingly contradicting requirement: $\rho$ should both change like a density and be invariant. The only way to satisfy both requirements is the following: particle states are fully identified by a set of coordinates $q^i$ and by conjugate set of variables $k_i$ whose units are the inverse of the ones defined by $q^i$. This way the product $dq^i dk_i$ is a pure number that is invariant under coordinate transformations. A coordinate transformation $\hat{q}^j = \hat{q}^j (q^i)$ induces a transformation $\hat{k}_j = \frac{\partial q^i }{\partial \hat{q}^j} k_i$ such that the area element $d\hat{q}^j d\hat{k}_j = dq^i \frac{\partial \hat{q}^j }{\partial q^i} \frac{\partial q^i }{\partial \hat{q}^j} dk_i=dq^i dk_i$ is invariant. An independent degree of freedom, then, is a pair $(q,k)$ for which the units can be changed independently from all other state variables. Mathematically, symplectic manifolds are the only spaces that allow us to define coordinate invariant densities. We can introduce a constant $\hbar$ such that $p_i = \hbar k_i$ which at this point is simply an arbitrary choice of unit.

These results gives us a precise understanding of what each layer of the mathematical structure corresponds to physically:
\begin{description}[noitemsep]
	\item \emph{Set of points:} possible states for the particles.
	\item \emph{Topological space:} ability to experimentally distinguish the states.
	\item \emph{Manifold:} states can be distinguished with a set of real valued quantities or state variables.
	\item \emph{Differentiable manifold:} we can define a distribution over the states and the state variables.
	\item \emph{Symplectic manifold:} we can define a distribution over the states that are coordinate independent.
\end{description}

\subsubsection{Hamiltonian mechanics}

\textbf{Determinism and reversibility.} A system undergoes deterministic and reversible evolution if given the initial state one can predict the final state and given the final state one can reconstruct the initial state. In other words, the system is predictable and retrodictable.

\emph{Reversibility as reconstructing the past.} This definition of reversibility should not be confused with time symmetry or the ability to undo the change. The first one means that if one substitute $t \to -t$ the dynamics is invariant (i.e. the same) or form invariant (i.e. different but described by the same equations) depending on the author. The second one means that we can find another process that brings the final state back to the original state. These concepts do not seem to work well in the general case. For example, when discussing reversibility in the sense of undoing the changes, one typically discusses changing the direction of the velocities to bring about the initial state with inverted velocities. In the presence of forces or an arbitrary dynamics, this operation because more complicated or ill-defined. The core problem with this definition is that it is not a property of the process itself, but on the existence of other processes (i.e. the process that undoes the changes or the process that flips the velocities, therefore changing the state). The requirement of time-invariance is too restrictive as it excludes all time dependent processes. Form invariance on an arbitrary set of equations, instead, can be broken by a suitably complicated change of variables, so it is only well-defined for a limited number of systems (e.g. particles under potential forces). The definition of reversibility we use is preferable because it works more generally and, when appropriate conditions apply, reduces to the other two which are better regarded as special cases. It is mathematically more elegant as it is the dual of determinism. It connects directly with information theory, as it states that the present has enough information to reconstruct the past, which gives a logical bridge to connect reversibility with information entropy and thermodynamics/statistical mechanics entropy.

\emph{Determinism and reversibility imply Hamiltonian mechanics.} In the context of infinitesimal reducibility, determinism and reversibility means that all the particles and only the particle within a given initial state $\rho(s(t)) = \rho(s(t+\Delta t))$ will be mapped to a particular final state. That is, the densities will be preserved over time. Time evolution, then, is a canonical transformation in phase space. As time is continuous, we can find a function $H : \mathcal{S} \to \mathbb{R}$ which acts as a time generator, which we recognize as being the Hamiltonian. This is essentially Louisville's theorem applied in reverse.

\emph{Invariant densities under time transformations imply extended phase-space.} So far, time has been treated as a parameter. If we introduce coordinate transformation that mix time, however, this is no longer possible. If $d\hat{q} = dq + vdt$ then the densities need to be defined on $dt$ as well. To do that, we have to introduce a conjugate variable $\omega$ of unit inverse time so that $dt d\omega$ is dimensionless and invariant. The invariant area element becomes $\sum_i dq^i dk_i - dt d\omega$ or $\sum_i dq^i dp_i - dt dE$ in units of $\hbar$. For surfaces at equal time $dt=0$ it reduces to the previous case. For surfaces that are not at equal time, since the temporal degree of freedom does cannot introduce new states, we need to correct the area element to account for the larger areas measured by the state variables that does not correspond to new states. If we set $q^\alpha = [t, q^i]$ and $p_\alpha = [-E, p_i]$ the area element becomes $dq^\alpha dp_\alpha$. These are elements of special relativity that come just by requiring invariance under time transformation. They do not come from invariance of the speed of light, they do not require a metric tensore. In other words, there is no way to have time transformations in Hamiltonian mechanics without requiring a four-momentum co-vector. Special relativity is not an independent choice: it is already partially baked into the two assumptions we took.

\emph{Extended phase-space imply classical anti-particles.} The trajectories in the extended phase space are in terms of an affine parameter $s$ and not time: $q^\alpha(s)$. As the motion is still deterministic and reversible, though, time $t(s)$ must be monotonic with respect to the affine parameter. For each state, the direction of time can be either aligned or anti-aligned with the affine parameter. Because the relationship is monotonic, aligned states can only connect to other aligned states and anti-aligned states can only connect to other anti-aligned states. We call aligned states the particle states and the anti-aligned states the anti-particle states.

\emph{Time evolution is better expressed by the Hamiltonian constraint.} The temporal degree of freedom introduces two variables but, since one is energy, there is an additional constraint since the value of the energy is linked to the other variables through the Hamiltonian. This link is better expressed by the Hamiltonian constraint $\mathcal{H} : \mathcal{\bar{S}} \to \mathbb{R}$ which is a function over the extended phase space, just like the Hamiltonian was a function over the standard phase space. The Hamiltonian constraint has two roles: it gives us the equation of motions in the same way of the standard Hamiltonian, and it constraints the motion over the surface $\mathcal{H} = 0$. To give context, the Hamiltonian constraint for a free particle is $\mathcal{H} = \frac{|p^\alpha|^2}{2m} - \frac{1}{2}mc^2$, which constraints the norm of the four-momentum to $mc$. The Hamiltonian constraint, unlike the standard Hamiltonian, is invariant under time transformation. It is the generator of the evolution in terms of an affine parameter.

\emph{Hamiltonian constraint as operator on distributions.} If $\rho$ is the distribution for the whole system, then $\mathcal{H} \rho = 0$. In fact, the distribution can only be non-zero on the constrain, for which the Hamiltonian constraint is zero. So, for each point in the extended phase space, either the distribution is zero or the Hamiltonian constrain is zero. Note that this is the same form one has in quantum mechanics for the Klein-Gordon and Dirac equations $\mathcal{H} |\psi\rangle = 0$, which in fact can be understood as Hamiltonian constraints that fix the norm of the momentum four-vector and four-velocity.

\emph{Determinism and reversibility is equivalent to conservation of information entropy.} Given a distribution, its information entropy is given by $I = - \int \rho \log \rho \, d\xi^n$. In general, information entropy is not invariant under change of variable: $- \int \rho(\hat{\xi}^b) \log \rho(\hat{\xi}^b) \, d\hat{\xi}^n = - \int \rho(\xi^a) \log \rho(\xi^a) \, d\xi^n - \int \rho(\xi^i) \log \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a}\right| \, d\xi^n$. Under a canonical transformation over phase space, though, the Jacobian determinant is unitary everyone therefore information entropy is invariant. That is, invariance of the density and invariance of its information entropy are one and the same. Requesting that densities are conserved in time is therefore equivalent to requesting that information entropy is conserved in time. This should be intuitive: the amount of information needed to identify an element within the initial or the final distribution is the same if the evolution is deterministic and reversible. Deterministic and reversible evolution, Hamiltonian evolution and conservation of information entropy, then, are one and the same.

\emph{Conservation of information entropy leads to classical uncertainty principle.} As information entropy is conserved, a Hamiltonian process can only map a distribution to one that has the same information entropy. If we consider a single degree of freedom $(q, p)$ and fix the information entropy $I_0$, then the distribution that minimizes the product of the standard deviation is the Gaussian. The information entropy for the Gaussian is given by $I_G = \ln (2\pi e \sigma_q \sigma_p)$. During the evolution, we have:
$$ \sigma_q \sigma_p \geq \frac{e^{I_0}}{2\pi e} $$.

\subsubsection{Lagrangian mechanics and massive particles}

\textbf{Kinematic equivalence.} A system is said to satisfy kinematic equivalence if giving the state of the system is equivalent to giving the trajectory in space and vice-versa. That is, for each possible state there is one and only one trajectory associated with it. Note that this is not true in general: for example, given the trajectory of a photon we cannot reconstruct its momentum since all photons travel at the same speed.

Note that we use $(q^\alpha, p_\alpha)$ for position and momentum and $(x^\alpha, u^\alpha)$ for position and velocity even though $q^\alpha = x^\alpha$. We do this because $\frac{\partial}{\partial q^\alpha} \neq \frac{\partial}{\partial x^\alpha}$ since one is taken at constant $p_\alpha$ and the other at constant $u^\alpha$. This is a tremendous source of confusion which is avoided by the use of the two symbols.

\emph{Kinematic equivalence means having an invertible relationship between velocity and momentum.} If the state of the system $(q^\alpha, p_\alpha)$ is enough to determine the trajectory $x^\alpha(s)$, then it will be enough to determine the position and velocity $(x^\alpha, u^\alpha)$. These will be enough to reconstruct the state as they span a space of equal dimension. As $q^\alpha = x^\alpha$, then $u^\alpha = u^\alpha(q^\beta, p_\gamma)$ is invertible.

\emph{Kinematic equivalence leads to Lagrangian mechanics.} Since $u^\alpha(q^\beta, p_\gamma)$, we can write $\mathcal{L} = \dot{q}^\alpha p_\alpha - \mathcal{H}$ as a function of $(x^\alpha, u^\alpha)$. That is the Lagrangian of the system. Note that this cannot be done if we do not have kinematic equivalence. For example, for a photon treated as a particle, we have $H = c |p|$. Using Hamilton's equations, we have $u^i = \frac{dx^i}{dt} = c \frac{p}{|p|}$ which is not invertible. While we can write the Lagrangian, this cannot be expressed in terms of position and velocity.

\emph{Kinematic equivalence leads to linear relationship between velocity and momentum.} Being able to go from $(q^\alpha, p_\alpha)$ to $(x^\alpha, u^\alpha)$ and vice-versa is not enough. We need to express the density $\rho$ in terms of position and velocity as well. This means expressing area elements $dq^\alpha dp_\alpha$ in terms of $dx^\alpha$ and $du^\alpha$. Since $dx^\alpha = dq^\alpha$, we must have $dp_\alpha = m g_{\alpha \beta} du^\beta$ where $g_{\alpha \beta}$ is an arbitrary linear function and $m$ is an arbitrary constant of proportionality.

\emph{Linear relationship between velocity and momentum leads to massive particles under potential forces.} We can integrate the relationship $dp_\alpha = m g_{\alpha \beta} du^\beta$ and find $p_\alpha = m g_{\alpha \beta} u^\beta + A_\alpha(x^\gamma)$ where $A_\alpha$ are arbitrary functions. Since $u^\alpha = \frac{dq^\alpha}{ds} = \frac{\partial \mathcal{H}}{\partial p_\alpha}$, we can take the previous relationship, integrate it again and find $\mathcal{H} = \frac{1}{2m} (p_{\alpha} - A_\alpha) g^{\alpha \beta} (p_{\beta} - A_\beta) + V$. We recognize this to be the Hamiltonian for a massive particle under potential forces.

\emph{Inertial mass measures the number of states per unit of velocity.} Inertial mass is usually introduced using the Newtonian relationship $F=ma$: the lower the mass, the easier is accelerate the body. If the mass where zero, then, we would expect the body to be infinitely easy to accelerate. If the mass is zero, however, the velocity is fixed and the body cannot be accelerated. The derivation offers a different insight: inertial mass is the constant of proportionality between ranges of velocity and ranges of states, as measured by momentum. The higher the mass, the more possible states are there per unit of velocity. If there are more states per unit of velocity, then the body is more difficult to accelerate as we have to go through more states, which recover the Newtonian intuition. However, if the mass is zero, set of states defined in a unit of velocity has measure zero: we cannot change state, we cannot accelerate.

\emph{The speed of light is a ration between the possible states in a unit of space over the possible states in a unit of time.} Regarding the constant $c$ as a speed is problematic. It is unclear why a speed, which is a property of motion, would be part of the metric of the space-time, which is defined regardless of motion and of what is there in the space itself. The derivation offers a different insight. As we move a state in space, it will go through different possible states: this can be quantified in phase space. As we move a state in time, it will also go through different possible states: this can be quantified on the extended phase space. The ratio between the two is the constant $c$. If we imagine a particle moving in both space and time, the position may remain the same but the time will always need to change. As the motion is deterministic and reversible, at most we can have one position in space for every position in time. Therefore, over a finite time, we must have $\frac{\Delta x}{\Delta t} \leq c$. Thinking about the constant $c$ as relating the count of states in space over time is more fundamental.

\emph{The Plank constant, the speed of light and the inertial mass serve to transfer the measure of states from one quantity to another.} As we saw initially, $(q, k)$ are the quantities on which we can measure the number of states. The constant $\hbar$ serves to transfer the count from $dk$ to $dp$. The inertial mass $m$ serves to transfer the count from $du$ to $dp$. The speed of light served to transfer the count from $dt$ to $dx$. They are book-keeping tools we use as we express states over different quantities which allow us to express in different ways the fundamental relationship coming from determinism and reversibility: the count of states must remain the same.

\subsubsection{Status and open issues}

We consider classical mechanics to be essentially forward engineered. There are some details, though, that could be better understood. The formalization will need a lot more time as it has many mathematical prerequisites.

\emph{Physical meaning of the Poisson brackets.} We still miss an understanding of the Poisson brackets as a generalization of Jacobian determinants. In principle, they should tells us how densities over a degree of freedom formed by two variables transform. We haven't found the right approach.

\emph{Densities as measure theoretic derivatives.} As we want finite objects to be our starting points, densities should be the ration of amount of ``stuff'' divided by the size of the region. This, in  measure theory, is known as the Radon-Nikodym derivative. We want this idea to be combined with statistics and differential geometry, so that we can talk about marginal distributions. It seems that gemoetric measure theory at least partially addresses that, which would be the first step for a full formalization of this section of the project.

\emph{Relationship between metric tensor and symplectic form.} The way the derivation works, the metric tensor is a secondary object that should be understood as being applied to a differential of position and a differential of velocity (e.g. $dx^\alpha g_{\alpha \beta} du^\beta$). This gives us the number of states on an area defined by the differentials. It is unclear whether the curvature of $g_{\alpha \beta}$ is linked to the geometry set by the symplectic form.

\subsection{Quantum mechanics}

\subsection{Thermodynamics}

\subsection{Statistical mechanics}

\section{The general theory}

Part of the project focuses on the development of a mathematical framework that provides the building blocks that are common to all scientific theories. That is, we need to formalize what is the minimum baseline that a theory needs to satisfy to be called a scientific theory.

This may seem a purely philosophical aspect, but in this project has actually a very pragmatic and technical implications. In any formal system, new structures are created by specializing existing ones. For example, when one writes a function or a class in a computer program, one is creating a specific instance with some extra feature, namely the functionality that one is implementing. This means that, as a baseline, we need to provide the general objects that one can specialize.

In mathematics, sets provides that general foundations. A group is a set with a set of operations that satisfy certain properties. A topology is a set together is a set of sets that satisfy certain properties. A vector space is a set together with some operations that satisfy certain properties. In mathematics, everything is a set with added special features. What we need to do is provide a similar foundation for science, where any scientific theory can be seen as a baseline plus some extra features.

\subsection{The Principle of Scientific Objectivity}

Our guiding idea is what we call the Principle of Scientific Objectivity, which states that ``Science is universal, non-contradictory and evidence based.'' Any scientific theory must be logically consistent, must be equally true or false for everybody and it must only talk about what can be established experimentally. Different branches of science will specialize on different topics (e.g. biology on living organisms, psychology on human behavior, chemistry on substances) but they all the theories and models they develop must satisfy those basic constraints.

Our principle indirectly gives us a definition for what we consider science and what it can study. There exist interesting subjects, such as the properties of prime numbers, moral questions or aesthetics of music, that are not universal (i.e. results may be subjective), contradictory (i.e. paradoxes may be unavoidable) or not-evidence based (i.e. answers not found through experimentation). These cannot be the object of science, and that is why we still need mathematics, arts, philosophy or religion.

It also tells us that only the part of nature that is experimentally accessible can be studied scientifically. For example, in control systems one defines a system as a state, a law of evolution, inputs and outputs. If one can use the inputs to reach any possible state, we say the system is controllable. If one can use the outputs to fully determines the state, we say the system is observable. When designing a control system, one ideally wants to have enough inputs and outputs to have an observable and controllable system. In the same vein, our ability to study a system scientifically depends on our ability to prepare and measure its properties reliably. This ability may change over time (i.e. thermodynamics developed after advances in temperature sensing)\footnote{It was the design of Otto von Guericke for the first vacuum pump that allowed Boyle and Hook to notice that the product of pressure times volume is constant. It was the invention of the mercury thermometer by Fahrenheit that allowed Gay-Lussac to notice that the pressure is proportional to the temperature. The ability to prepare a gas at different pressure together with the ability to measure the temperature of a gas precisely is what gave us the law of an ideal gas PV=nRT.} or may work only under certain circumstances (i.e. temperature is well defined only at equilibrium).

A scientific theory, then, is an objective non-contradictory model about what, under a set of circumstances, is experimentally accessible. Some may argue that the true objective of science is to give a full description of reality; or that scientific theories only deal with the mere knowledge we have of a system. We are not going to argue these positions. We do note that they are in contradiction with the principle of scientific objectivity: knowledge is not objective and there is nothing that tells us that all of reality is experimentally accessible. We also note that from this premise we can recover the correct mathematical structures needed to rederive the known theories. We take that as confirmation that our premise is valid.

\subsection{The Logic of Verifiable statements}

The basic building block for a scientific theory is the idea of a verifiable statement: an assertion that is either true or false for everybody and for which we have an experimental test that will terminate successfully in finite time if and only if the statement is true. What we need to do is augment standard logic with experimental verifiability.

If a scientific theory is a model, then it is a collection of statements with well-defined logical relationships between them. For example, a model of the solar systems will be able to say characterized by statements such as ``the earth rotates around

* Formal theories are specializations of more general ones
* Algebraic structures vs Propositional structures
* Physical theories as models; the theory of models
* Difference from ontological theories and subjective theories; accessible information
* Verifiable statements
* Meaning of statements can't be captured, requires context

\end{document}

### 2019-11-26 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
Title change
Adding Julian comments

@@
 
-\title{Universal Characterization of Physical Variability by \\Shannon's $-\sum_ip_i\,\textrm{ln}\,p_i$ Formula}
+\title{Variability as a better characterization of Shannon entropy}
 
@@
 
-%Cite Vlatko Vedral: Decoding Reality: The Universe as Quantum Information. Information is "the fundamental building block of the Universe".
-
-%Cite Paul Davies Information and the Nature of Reality. Information "occupies the ontological basement"
-
-%Cite Wheeler It from bit.
-"It is wrong, moreover, to regard this or that  physical quantity as sitting "out there" with this or that numerical value it is absorbed." "The information thus solicited make physics and comes in bits"
-
 We have seen what the Shannon variability measures and why it is important in information theory, but we have so far worked with discrete quantities.
@@
 \begin{itemize}
-	\item the Shannon entropy measures the variability of the elements within a given distribution, giving it a crisp intuitive meaning that is a general and applicable to all branches of science
-	\item the expression is not arbitrary, as it is the only linear indicator for such concept
+	\item the Shannon entropy measures the variability of the elements within a given distribution, giving it a crisp intuitive meaning that is general and applicable to all branches of science
+	\item the expression is not arbitrary, as it is the only linear indicator for such a concept
 	\item it measures the variability by quantifying the number of yes/no questions one must ask to identify an element within the distribution, which corresponds to the number of bits needed to transmit or store that information

### 2019-11-26 Articles/Overview/overview.tex
Starting to work on an overview of the project

@@
 The aim of this article is to give a broad overview of the project ``Assumptions of Physics'' so that the reader will understand its motivations, its status and a list of the main findings.
-This is intended to be a good start for anyone interested in collaborating or simply learning more about the project.
+This is intended to be a good start for anyone interested in simply learning more about the project or finding an area for collaboration.
+
+As we plan to update this document, please make sure you have a recent version.
 Each subsection will be fairly self-contained to allow the reader to skim through the document and focus on the parts of interest.
-
-As we plan to update this document, please make sure this is a recent version.
+It will contain a summary section at the beginning and a titled paragraph for each of the main idea or finding.
 
@@
 
+\section{The general mathematical theory of experimental science}
+
+This part of the project focuses on the development of a rigorous mathematical framework that can provide the building blocks that are common to all scientific theories.
+As logic and set theory provides the foundation for all other mathematical structure, which are essentially sets with operations defined on them, the general theory needs to provide the same fundamental layer.
+
+\subsection{The Principle of Scientific Objectivity}
+
+Our guiding is that ``Science is universal, non-contradictory and evidence based.'' Any scientific theory must be logically consistent, its content be equally true or false for everybody and it must deal with what can be established experimentally.
+Different branches of science will specialize on different systems and topics, but they all the theories and models they develop must satisfy those basic constraints.
+The general theory deals with the fundamental mathematical structure that realize that principle.
+
+\emph{Not every subject can be studied by science.} The principle indirectly states that anything that is subjective, contradictory or not evidence based cannot be the subject of scientific investigation.
+The properties of prime numbers, moral and existential questions or the aesthetics of music fail in at least on of those respects and are excluded.
+Therefore science cannot be a substitute of mathematics, arts, philosophy or religion.
+Only the parts of nature that are accessible through consistent experimental verification can be studied scientifically.
+New experimental techniques extend its reach.
+
+\subsection{The Logic of Verifiable Statements}
+
+The basic building block in our framework is the idea of a verifiable statement: an assertion that is either true or false for everybody and for which we have an experimental test that will terminate successfully in finite time if and only if the statement is true.
+The basic structures of our general theory need to provide a logic framework to keep track of verifiable statements and their relationships.
+
+\subsubsection{Logical contexts}
+
+A logical context consists of a set of statements with well defined logical relationships.
+It is the most fundamental structure and it is the only one that is axiomatically introduced.
+A scientific theory will consist of a set of statements taken within a logical context with particular logical relationships.
+For example, Newtonian mechanics will use statements like ``the mass of the object is 1 $Kg$'' and ``the acceleration of the object is 1 $m/s^2$''.
+If both of these are true, then the statement ``the force on the object is  1 $N$'' is also true.
+The context defines basic logical relationships (e.g.
+equivalence, narrowness, compatibility and independence) which form the basis for higher level constructs (e.g.
+ordering, linear and statistical independence).
+Every other structure imposed on the statements will need to ``play'' nice with these fundamental relationships.
+
+\emph{Note on terminology.} The terminology may depart slightly from what logicians may be accustomed to, simply because our aim is different.
+Namely, our goal is not to study the rule of inference or study what can be formally proven, but simply to keep track of the logical relationships between statements that are assumed to be already given.
+
+\emph{Statements as primary object.} Every formal theory needs prime objects, elements that will not further be described by the theory.
+We will take statements, assertion that are either true or false for everybody, as these objects.
+The formal system will specifically:
+\begin{description}
+	\item not be propositional; the language of the statements, its syntax, its grammar, will be left unspecified; statements will represent the content regardless of how it is expressed
+	\item be algebraic; operations on statements will not form new statements, but describe relationships within a given set of statements; note that most mathematical structures used in science are algebraic (computer science is a notable exception)
+	\item purely formal; it will not not try to capture the semantic of the statements, though the semantic will impose logical relationships that will be formally captured
+\end{description}
+
+\emph{Statements cannot exist in isolation.} Statements should, in general, not be regarded as independent of the context they are part of.
+Take the statement ``The mass of the electron is 510 $\pm$ 0.5 KeV''.
+If we are measuring the mass of the electron, that is a statement that may or may not be true depending on experiment.
+If we are performing particle identification in a detector, that statement is assumed to be true.
+The truth values allowed by the same statement depends on context.
+
+\emph{Logical context as the fundamental structure.} A logical context is a set of statements with well defined logical relationships.
+It defines which truth assignments are possible (i.e.
+are consistent with the semantic of the statements).
+It identifies one of the possible assignments as the true one.
+It allows to find statements whose truth value depends on others through a particular Boolean function.
+The existence of this structure is introduced by axioms as they are crossing the lines between physical objects and mathematical ones.
+
+\emph{Tautologies, contradictions and contingent statements.} We define tautologies as those statements that are true in every assignment, contradictions as those statements that are false in every assignment and contingent all the remaining.
+
+\emph{Statement equivalence and Boolean algebra.} Within our formal system, we can distinguish between:
+\begin{description}
+	\item statement equality: two statements that re-express the same fact in different languages or in different words
+	\item statement equivalence: two statements have the same truth value in all possible assignments
+	\item statement material equivalence: two statements have the same truth value
+\end{description}
+A logical context is a complete Boolean algebra under statement equivalence.
+
+\emph{Narrowness and ordering.} A statement is narrower then another if whenever it is true then the other is also true.
+Narrowness can often be thought as implication, except in a few corner case (e.g.
+a statement that is always false is narrower than any other statement).
+Narrowness imposes a partial order to the structure.
+This is the same order associated to the fact that every Boolean algebra is a complemented distributed lattice.
+
+\emph{Compatibility and independence.} Two statements are said to be compatible if they can both be true in the same assignment.
+Two statements are independent if fixing the truth value of one does not restrict the truth value of the other.
+Independence, like linear independence or statistical independence of which it is the foundation, is a property of a set of statements and is not transitive.
+
+\subsubsection{Verifiable statements}
+
+Logical contexts need to keep track which statements are experimentally verifiable.
+A statement is experimentally verifiable if we are provided with a test that, if the statement is true, will always terminate successfully in a finite time.
+The fact that the test has to terminate in a finite time in the positive case and that it may not terminate in the negative case has profound implications: verifiable statements are not closed under the standard logical connectors.
+The negation of a verifiable statement is not in, general, a verifiable statement.
+The finite conjuntion (i.e.
+logical AND) of verifiable statements is a verifiable statement, but not an infinite one.
+The countable disjunction (i.e.
+logical OR) of verifiable statements is a verifiable statement, but not a more then countable infinite one.
+
+\emph{TODO.} The negation of a verifiable statement is not in general a verifiable statement.
+We may be able to verify that ``there exists extra-terrestrial life'' if at some point we find a clear signal.
+Absence of signal, however, will never confirm that ``there is no extra-terrestrial life''.
+
+\subsubsection{Experimental domains}
+
+We will define an experimental domain as a set of verifiable statements that can be expressed as the combination of a countable subset (i.e.
+a countable base).
+Because of finite time verifiability, this is the biggest space we can experimentally probe given an indefinite amount of time.
+An experimental domain, then, represents all the information that can be gathered experimentally about a particular subject.
+
+From each experimental domain we construct a theoretical domain by allowing negation as well.
+This will include all statements for which an experimental test is in principle possible, though there is no guarantee of termination.
+A theoretical domain, then, represents all statement to which we can attach a prediction.
+
+Within a theoretical domain we define the set of possibilities as those statements that, once known to be true, will set the truth value of all other statements.
+Each possibility represents a possible case that is distinguishable experimentally, which corresponds to a unique possible assignment for the experimental and theoretical domain.
+Because of the countable base, the set of possibilities can never be greater than the continuum.
+
+To sum up, an experimental domain consists of a set of statements that can be tested experimentally (e.g.
+``the animal has whiskers'', ``the mass of the photon is less than $10^{-13} eV$'').
+The theoretical domains extends to statements that may not be tested experimentally (e.g.
+``the mass of the photon is exactly 0 eV'').
+The possibilities consists of all the possible cases (e.g.
+``the animal is a cat'', ``the mass of the photon is exactly $0.102 10^-{34} eV$ ).
+
+\subsubsection{Topologies and $\sigma$-algebras}
+
+An experimental domains provide a natural topology over the possibilities.
+Each verifiable statements can be identified with the set of possibilities compatible with it (e.g.
+the statement ``the mass of the photon is less than $10^{-13} eV$'' is equal to the disjuntion of all statement of the form ``the mass of the photon is exactly x $eV$'' with $x <  10^{-13}$).
+Since verifiable statements are closed under finite conjunction and countable disjunction, the sets corresponding to verifiable statements form a $T_0$ second countable topology.
+
+On the other hand, theoretical domains provide a natural $\sigma$-algebra over the possibilities.
+Each theorical statement can also be identified with the set of possibilities compatible with it.
+Since theoretical statements are closed under negation and countable disjunction, the sets corresponding to theoretical statements form a $\sigma$-algebra, which is the Borel algebra of the natural topology.
+
+Topologies and $\sigma$-algebras provide the foundations for differential geometry, Lie algebras, measure theory, probability theory and many other mathematical tools used in physics and the sciences.
+As we have now have a precise understanding of what they represent, all concepts and proofs in those subject can be understood in terms of experimental verifiability.
+
+\subsubsection{Status and open issues}
+
+This part of the work is very well developed.
+More work could be done in finding meaning for all mathematical concepts (e.g.
+compact sets, all separability axioms, ...).
+It would be interesting to work with people from different fields (i.e.
+logic, foundations of mathematics, philosophy of science, ...) to see whether any aspect of this section would appeal to different communities.
+
+\subsection{Common constructions}
+
+With the basic mathematical structure defined, we develop some definitions and constructions that are of general use.
+
+\subsubsection{Relationships and equivalence between domains}
+
+We define two relationships between domain: inference relationships and causal relationships.
+An inference relationship establishes that testing a verifiable statement in one domain is the same as testing a verifiable statement in the other.
+Mathematically it is a map that takes a verifiable statement from the one and returns a verifiable statement from the other that is equivalent to the first.
+A causal relationship establishes that determining which possibility is true in one domain also determines which possibility is true in the second.
+Mathematically it is a map that given a possibility of the first returns a possibility of the second that is broader than (i.e.
+implied from) the first.
+
+Among the main results we find that there exists an inference implication between two domains if and only if there exists a causal relationship between them.
+The direction of the inference is the opposite of the causal direction.
+Therefore we say that a domain depends on another if there exists an inference relationship between the first and the second or if there exists a causal relationship between the second and the first.
+Causal relationships must be continuous function in terms of the natural topologies, which justifies why functions in science are always assumed to be ``well-behaved'' (i.e.
+analytically continuous with at most countable discontinuities).
+
+\subsubsection{Composite domains}
+
+\subsubsection{Relationship domains}
+
+\subsubsection{Status and open issues}
+
+The overall mechanics of composite domains and relationships domains are fairly well developed and understood.
+There are details, though, that can be understood better.
+The connection to category theory can be also better developed.
+Other common concepts and constructions, like similarities or sequences of domains, could be formulated.
+
+\subsection{Properties and Quantities}
+
+\subsubsection{Properties and quantities}
+
+\subsubsection{Ordering theorems for discrete and continuous quantities}
+
+If a scientific theory is a model, then it is a collection of statements with well-defined logical relationships between them.
+For example, a model of the solar systems will be able to say characterized by statements such as ``the earth rotates around
+
+* Formal theories are specializations of more general ones
+* Algebraic structures vs Propositional structures
+* Physical theories as models; the theory of models
+* Difference from ontological theories and subjective theories; accessible information
+* Verifiable statements
+* Meaning of statements can't be captured, requires context
+
 \section{Physics theories}
@@
 
-Classical Hamiltonian and Lagrangian particle mechanics, including the Hamiltonian for massive particles under potential forces, can be derived from three main assumptions: infinitesimal reducibility, determinism and reversible evolution and kinematic equivalence.
-We list here the main points of the derivation together with the some of the insights and connections it provides.
+Classical mechanics can be derived from three main assumptions: infinitesimal reducibility, determinism and reversible evolution and kinematic equivalence.
+From the first assumption we can show that the state space of the infinitesimal parts of a system (i.e.
+particles) has the structure of phase space.
+From the second we can show that the time evolution follows Hamiltonian mechanics.
+From the third we can show that the system is Lagranian and is restricted to the case of massive particles under potential forces.
 
@@
 We call particle an infinitesimal part of the system, the limit of recursive subdivision.
+The system is then described by a distribution of the state its parts over phase space, which is the only structure that allows coordinate invariant distributions over continuous variables.
 
@@
 In other words, the system is predictable and retrodictable.
+Densities at each past state must be equal to the density of the corresponding future state, which leads to Hamiltonian dynamics.
 
@@
 Note that this is not true in general: for example, given the trajectory of a photon we cannot reconstruct its momentum since all photons travel at the same speed.
+Under this assumption the relationship between position/momentum and position/velocity is invertible and the system admits a Lagrangian.
+Moreover, given the transformation rules of momentum and velocity, we find that there must be a linear relationship between the two, which constrains the dynamics to the one of massive particle under potential forces.
 
@@
 
-\section{The general theory}
-
-Part of the project focuses on the development of a mathematical framework that provides the building blocks that are common to all scientific theories.
-That is, we need to formalize what is the minimum baseline that a theory needs to satisfy to be called a scientific theory.
-
-This may seem a purely philosophical aspect, but in this project has actually a very pragmatic and technical implications.
-In any formal system, new structures are created by specializing existing ones.
-For example, when one writes a function or a class in a computer program, one is creating a specific instance with some extra feature, namely the functionality that one is implementing.
-This means that, as a baseline, we need to provide the general objects that one can specialize.
-
-In mathematics, sets provides that general foundations.
-A group is a set with a set of operations that satisfy certain properties.
-A topology is a set together is a set of sets that satisfy certain properties.
-A vector space is a set together with some operations that satisfy certain properties.
-In mathematics, everything is a set with added special features.
-What we need to do is provide a similar foundation for science, where any scientific theory can be seen as a baseline plus some extra features.
-
-\subsection{The Principle of Scientific Objectivity}
-
-Our guiding idea is what we call the Principle of Scientific Objectivity, which states that ``Science is universal, non-contradictory and evidence based.'' Any scientific theory must be logically consistent, must be equally true or false for everybody and it must only talk about what can be established experimentally.
-Different branches of science will specialize on different topics (e.g.
-biology on living organisms, psychology on human behavior, chemistry on substances) but they all the theories and models they develop must satisfy those basic constraints.
-
-Our principle indirectly gives us a definition for what we consider science and what it can study.
-There exist interesting subjects, such as the properties of prime numbers, moral questions or aesthetics of music, that are not universal (i.e.
-results may be subjective), contradictory (i.e.
-paradoxes may be unavoidable) or not-evidence based (i.e.
-answers not found through experimentation).
-These cannot be the object of science, and that is why we still need mathematics, arts, philosophy or religion.
-
-It also tells us that only the part of nature that is experimentally accessible can be studied scientifically.
-For example, in control systems one defines a system as a state, a law of evolution, inputs and outputs.
-If one can use the inputs to reach any possible state, we say the system is controllable.
-If one can use the outputs to fully determines the state, we say the system is observable.
-When designing a control system, one ideally wants to have enough inputs and outputs to have an observable and controllable system.
-In the same vein, our ability to study a system scientifically depends on our ability to prepare and measure its properties reliably.
-This ability may change over time (i.e.
-thermodynamics developed after advances in temperature sensing)\footnote{It was the design of Otto von Guericke for the first vacuum pump that allowed Boyle and Hook to notice that the product of pressure times volume is constant.
-It was the invention of the mercury thermometer by Fahrenheit that allowed Gay-Lussac to notice that the pressure is proportional to the temperature.
-The ability to prepare a gas at different pressure together with the ability to measure the temperature of a gas precisely is what gave us the law of an ideal gas PV=nRT.} or may work only under certain circumstances (i.e.
-temperature is well defined only at equilibrium).
-
-A scientific theory, then, is an objective non-contradictory model about what, under a set of circumstances, is experimentally accessible.
-Some may argue that the true objective of science is to give a full description of reality; or that scientific theories only deal with the mere knowledge we have of a system.
-We are not going to argue these positions.
-We do note that they are in contradiction with the principle of scientific objectivity: knowledge is not objective and there is nothing that tells us that all of reality is experimentally accessible.
-We also note that from this premise we can recover the correct mathematical structures needed to rederive the known theories.
-We take that as confirmation that our premise is valid.
-
-\subsection{The Logic of Verifiable statements}
-
-The basic building block for a scientific theory is the idea of a verifiable statement: an assertion that is either true or false for everybody and for which we have an experimental test that will terminate successfully in finite time if and only if the statement is true.
-What we need to do is augment standard logic with experimental verifiability.
-
-If a scientific theory is a model, then it is a collection of statements with well-defined logical relationships between them.
-For example, a model of the solar systems will be able to say characterized by statements such as ``the earth rotates around
-
-* Formal theories are specializations of more general ones
-* Algebraic structures vs Propositional structures
-* Physical theories as models; the theory of models
-* Difference from ontological theories and subjective theories; accessible information
-* Verifiable statements
-* Meaning of statements can't be captured, requires context
-
 \end{document}

### 2019-11-26 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
Adding affiliations for AJP submission

@@
-\documentclass[prb, twocolumn]{revtex4-1}
+\documentclass[prb,preprint]{revtex4-1} 
+% The line above defines the type of LaTeX document.
+% Note that AJP uses the same style as Phys. Rev. B (prb).
 
-%\documentclass{article}
-%\textheight 8.9 in \textwidth 6.5 in \oddsidemargin -10pt \topmargin -30pt
+% The % character begins a comment, which continues to the end of the line.
 
-\usepackage{amsmath}
-%\usepackage[usenames,dvipsnames]{color}
+\usepackage{amsmath}  % needed for \tfrac, \bmatrix, etc.
+\usepackage{amsfonts} % needed for bold Greek, Fraktur, and blackboard bold
+\usepackage{graphicx} % needed for figures
 
@@
 \author{Gabriele Carcassi}
+\email{carcassi@umich.edu}
+\author{Christine A. Aidala}
+\email{caidala@umich.edu}
+\affiliation{Department of Physics, University of Michigan, Ann Arbor, MI 48109-1040 }
+% Please provide a full mailing address here.
+
 \author{Julian Barbour}
-\author{Christine A. Aidala}
+\altaffiliation[permanent address: ]{College Farm, South Newington, Banbury, Oxon, OX15 4JG UK}
+\affiliation{Visiting Professor in Physics at the University of Oxford, UK}
+
+\date{\today}
 

### 2019-11-30 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
Final touches before submission

@@
 \author{Julian Barbour}
+\email{julian.barbour@physics.ox.ac.uk}
 \altaffiliation[permanent address: ]{College Farm, South Newington, Banbury, Oxon, OX15 4JG UK}
-\affiliation{Visiting Professor in Physics at the University of Oxford, UK}
+\affiliation{Visiting Professor in Physics at the University of Oxford (2009-2018)}
 
@@
 
-The Shannon entropy $-\sum p_i \log p_i$ is well established in such disparate fields as computer science[cite], physics[cite] and ecology[cite].
-It is most often introduced in physics through statistical mechanics, and associated with vague notions, such as uncertainty[cite], knowledge[cite], lack of knowledge[cite] and disorder[cite].
+The Shannon entropy $-\sum p_i \log p_i$ is well established in such disparate fields as computer science, physics and ecology.
+It is most often introduced in physics through statistical mechanics, and associated with vague notions, such as uncertainty, knowledge, lack of knowledge and disorder.
 These prevent a proper understanding of the concept and often lead to confusion.
@@
 We conclude noting that in quantum mechanics the Shannon variability, in the form of the von Neumann entropy, is coordinate invariant as well.
-That is, given a density matrix $\rho$ then $S_{VN} = - \textrm{tr}(\rho \log \rho)$ is independent of the basis in which it is calculated.
+That is, given a density matrix $\rho$ then $S_\textrm{\tiny{vN}} = - \textrm{tr}(\rho \log \rho)$ is independent of the basis in which it is calculated.
 
@@
 If $N$ is large, the number of particles in one region $N_\Sigma$ is very close to the expectation value $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$.
-If these assumptions are not met, we cannot break the joint distribution into single-particle ones, particle number and type may change, the fluctuations may become relevant thus requiring a more general account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.}
+If these assumptions are not met, we cannot break the joint distribution into single-particle ones, particle number and type may change, the fluctuations may become relevant thus requiring a more general account.\footnote{Jaynes\cite{Jaynes} has shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.}
 
@@
 The first is that $\Gamma$-space automatically assumes that all particles are distinguishable.
-This leads to the widely known problem of overcounting which needs to be addressed in the standard way.[cite] The second problem is that, though the state of each particle is given by position and momentum, we should not think of them as literally pointlike.
+This leads to the widely known problem of overcounting which needs to be addressed in the standard way.
+The second problem is that, though the state of each particle is given by position and momentum, we should not think of them as literally pointlike.
 As we said, this would correspond to delta Dirac distributions over phase space which have minus infinite Shannon variability.
@@
 We note that some authors choose to interpret the probability distribution not as coming from repeated independent trials, but as the knowledge one has about the system.
-This would make the entropy a subjective notion\footnote{Jaynes called it the anthropomorphic nature of entropy.}, but the associated fluctuations we experimentally observe are indeed objective.
-We believe the confusion comes from the, correct, realization that the same system, under different conditions, will be described with a different set of thermodynamic variables.\footnote{Jaynes points out:
+This would make the entropy a subjective notion\footnote{Jaynes\cite{Jaynes} called it the anthropomorphic nature of entropy.}, but the associated fluctuations we experimentally observe are indeed objective.
+We believe the confusion comes from the, correct, realization that the same system, under different conditions, will be described with a different set of thermodynamic variables.\footnote{Jaynes\cite{Jaynes} points out:
 \begin{quote}
@@
 Instead of a distribution over the $N$-particle phase space, we have a distribution over the Hilbert space for the quantum system which is represented by a density matrix operator $\rho$.
-The von Neumann entropy is given by $S_{V\!N} = - \textrm{tr}(\rho \log \rho)$.
-This, expanded in a basis, becomes $S_{V\!N} = - \sum_i \rho_i \log \rho_i$ or $S_{V\!N} = - \int \rho(x) \log \rho(x)dx$ depending on whether the spectrum is discrete or continuous.
+The von Neumann entropy is given by $S_\textrm{\tiny{vN}} = - \textrm{tr}(\rho \log \rho)$.
+This, expanded in a basis, becomes $S_\textrm{\tiny{vN}} = - \sum_i \rho_i \log \rho_i$ or $S_\textrm{\tiny{vN}} = - \int \rho(x) \log \rho(x)dx$ depending on whether the spectrum is discrete or continuous.
 The von Neumann entropy corresponds to the Shannon variability of the microstate distribution for a given quantum macrostate.

### 2019-12-13 Articles/Overview/overview.tex
More work on the overview

@@
 
-The aim of this article is to give a broad overview of the project ``Assumptions of Physics'' so that the reader will understand its motivations, its status and a list of the main findings.
-This is intended to be a good start for anyone interested in simply learning more about the project or finding an area for collaboration.
+The aim of this article is to give a broad overview of the project ``Assumptions of Physics'': its motivations, its status and a list of the main findings.
+It is intended to be a good start for anyone interested in simply learning more about the project or finding an area for collaboration.
 
@@
 
-A formal framework that forces us to clarify our assumptions
-From those assumptions the mathematical objects are derived
-Each mathematical object has a clear physical meaning and no object is unphysical
-Gives us concepts and tools that span across different disciplines
-Allows to explore what happens when the assumptions fail, possibly leading to new physics ideas
-
 \textbf{A better understanding of physics.} The overall goal of the project is to better understand physics: to understand why the laws of quantum mechanics, or classical mechanics, are what they are; to understand the limit of the validity of the different theories; to understand what the Poisson brackets, or any other mathematical operation, represent physically; to understand which mathematical objects correspond to actual physical entities and which are just artifacts; to understand what new theories are possible.
 
-To us, understanding means to see the necessity of something and how it relates to the rest.
-For example, once we know that the earth's revolution axis is not aligned with earth's rotation axis, we understand why there must be seasons; and we understand why days are shorter in the winter; and why the sun is brighter and warmer in the summer.
-The more connections, the fewer starting points, the more understanding.
+Ideally, we want to achieve a single unified framework for all science that is physically meaningful (i.e.
+it is clear what each objects represents experimentally and why must it be represented that way), mathematically precise (i.e.
+the physical assumptions are captured through axioms and definitions and then the arguments are carried out with the rigor of modern mathematics) and philosophically consistent (i.e.
+the concepts and viewpoints taken fit with the mathematical results and the practice of science).
 
@@
 
-\textbf{Give science sturdier mathematical grounds.} Physicists are generally trained to think that mathematical details are physically not interesting and are the domain of mathematicians.
-Even mathematical structures that originally came from physics are now formalized by mathematicians to solve their own needs, and not the physicist's.
+\textbf{Give science sturdier mathematical grounds.} There is a general sense within physics that, at some point, mathematical details are physically not interesting and should be handled ``appropriately'' by mathematicians.
+Therefore even mathematical structures that originally came from physics are now formalized by mathematicians to solve their own needs, and not the physicist's.
 This, in turns, reinforces the idea that those details are not interesting to the physicists.
-The reality is that a well-posed physical problem must also be a well-posed mathematical problem, but not the other way around, so leaving the details to the mathematicians means having fundamental structures that are less physical.
-For example, quantum states are modeled with Hilbert spaces even though we know they contain mathematical objects that are unphysical.
+The reality is that a well-posed physical problem must also be a well-posed mathematical problem, but not the other way around, so leaving the details to the mathematicians means having fundamental structures that are less physical.\footnote{For example, quantum states are modeled with Hilbert spaces even though we know they contain mathematical objects that are unphysical.} Mathematicians will come up with definitions so that theorems are easier to prove or calculations are easier to perform.
+They will not come up with definitions based on whether they are physically meaningful.
 If we give a theory, as we said before, by simply stating a mathematical structure, then there is no guarantee that what we give is a fully physical theory.
@@
 \textbf{Provide a solid basis for new theories.} Most of the attention in fundamental physics is focused on the development of new theory or the search for new effects.
-A better understanding of the current theories, recast in a single broader and more precise framework, facilitates that search.
-It gives more insight and intuition leading to new ideas.
-It makes explicit what ingredients went into the theories, so we know what assumption can fail and what principles cannot be changed.
-For example, we may abandon the assumption that space is measurable at an indefinite precision, leaving the realm of manifold and real numbers; yet we cannot abandon the notion that space is measurable, staying in the realm of second countable $T_0$ topological spaces.
-As all the physical theories are part of a more general structure, they also provide templates for new theories and prove that the more general structure is sound.
+A better understanding of the current theories, recast in a single broader and more precise framework, will probably facilitate that search.
+It makes explicit what ingredients went into the theories, so we know what assumption can fail and what principles cannot be changed.\footnote{For example, we may abandon the assumption that space is measurable at an indefinite precision, leaving the realm of manifold and real numbers; yet we cannot abandon the notion that space is experimentally accessible, staying in the realm of second countable $T_0$ topological spaces.} If all physical theories can be seen as instances of a more general structure, they prove that the general structure is sound and provide templates for new theories.
 
-The objectives are cover many areas and are very broad.
-As such, we are always looking for people with diverse background and interest.
-We ultimately believe that all these objectives are connected and cannot be pursued independently.
+While the objectives cover many areas and are very broad, our experience tells us that they are ultimately connected and cannot be pursued to their fullest independently.
+Therefore we are always looking for people with diverse background and interest to help us cover the different areas.
 
@@
 \textbf{Reverse engineering.} The goal of this phase is to deconstruct an already existing theory or mathematical structure to understand what each piece is supposed to represent physically.
-One way to conduct such investigation is to ask what would happen if a particular characteristic of the framework would fail.
-For example, a line of reasoning could be as follows: Hamilton's equations are differential; if they weren't the mapping of densities and areas would not be well defined; maybe Hamiltonian mechanics is not about mapping points but about mapping densities or areas.
-This usually leads to a set of necessary assumptions about the physical system one is describing, the ones one must take for the particular description to apply.
+One way to conduct such investigation is to ask what would happen if a particular characteristic of the framework would fail.\footnote{For example, a line of reasoning could be as follows: Hamilton's equations are differential; if they weren't the mapping of densities and areas would not be well defined; maybe Hamiltonian mechanics is not about mapping points but about mapping densities or areas.} This usually leads to a set of necessary assumptions about the physical system one is describing, the ones one must take for the particular description to apply.
 
 \textbf{Forward engineering.} The goal of this phase is to construct an argument from the necessary assumptions outlined in the previous phase to rederive the framework at hand.
-This checks whether the necessary assumptions are also sufficient.
-For example, if we started simply by assuming we have density distributions mapped in time, this would give us all possible differentiable evolutions, not necessarily Hamiltonian ones.
-This leads to a set of necessary and sufficient conditions to rederive the theory.
+This checks whether the necessary assumptions are also sufficient \footnote{For example, if we started simply by assuming we have density distributions mapped in time, this would give us all possible differentiable evolutions, not necessarily Hamiltonian ones.}.
+It ends when a set of necessary and sufficient conditions to rederive the theory are found, meaning that the full theory has been physically characterized.
 
@@
 This forces us to clarify all the starting points and what parts of the arguments are truly formal and what are physical.
-This can be done only starting from the foundations and therefore takes a considerable longer time.
-For example, before formalizing Hamiltonian mechanics we have to formalize differentiable manifolds, real number and topological spaces.
+Unfortunately, this can ultimately be done only by starting from the very foundations\footnote{For example, before formalizing Hamiltonian mechanics we have to formalize differentiable manifolds, real numbers and topological spaces.} though we are at a point where an overall tentative structure is emerging and tentative formalization can be attempted on higher structures.
 
@@
 \textbf{The general theory.} This formalizes the basic mathematical framework that is the basis for all science.
-Like in mathematics every algebraic structure is a specialization of the generic structures axiomatized by set theory, all scientific theory are ultimately specialization of the generic structure provided by the general theory.
+As every algebraic structure in mathematics is a specialization of the generic structures axiomatized by set theory, all scientific theory are ultimately specialization of the generic structure provided by the general theory.
 This defines the basic requirements imposed on a theory by logical consistency and experimental verifiability.
@@
 
-\textbf{Fundamental physics concepts.} This defines the notions that are fundamental to physics specifically, such as states, processes, environment, equilibrium, determinism and so on.
-The idea is to define these concepts in the most abstract way possible such that common requirements can be identified.
+\textbf{Physics core.} This defines the notions that are fundamental to physics specifically, such as states, processes, environment, equilibrium, determinism and so on.
+The idea is to define these concepts in the most general way possible such that common requirements can be identified.
 Every physical theory is a further refinement of this structure in that it will choose to study a particular set of states over a particular process.
@@
 
-We will present the work in the reverse order, from physical theories to general theory, as we find most physicist
-
-\section{The general mathematical theory of experimental science}
+\section{The general theory}
 
 This part of the project focuses on the development of a rigorous mathematical framework that can provide the building blocks that are common to all scientific theories.
-As logic and set theory provides the foundation for all other mathematical structure, which are essentially sets with operations defined on them, the general theory needs to provide the same fundamental layer.
+As logic and set theory provides the foundation for all other mathematical structures, which are essentially sets with associated operations, the general theory will define common tools (e.g.
+logic of experimental verification, causal relationships, measurable quantities) that other theories will specialize in different ways.
 
@@
 
-Our guiding is that ``Science is universal, non-contradictory and evidence based.'' Any scientific theory must be logically consistent, its content be equally true or false for everybody and it must deal with what can be established experimentally.
-Different branches of science will specialize on different systems and topics, but they all the theories and models they develop must satisfy those basic constraints.
-The general theory deals with the fundamental mathematical structure that realize that principle.
+Our guiding principle is that ``Science is universal, non-contradictory and evidence based.'' Any scientific theory must be logically consistent, its content be equally true or false for everybody and it must deal with what can be established experimentally.
+Different branches of science will specialize on different systems and topics, but all the theories and models they develop must satisfy those basic constraints.
+The general theory deals with the fundamental mathematical structure needed to realize that principle.
 
 \emph{Not every subject can be studied by science.} The principle indirectly states that anything that is subjective, contradictory or not evidence based cannot be the subject of scientific investigation.
-The properties of prime numbers, moral and existential questions or the aesthetics of music fail in at least on of those respects and are excluded.
-Therefore science cannot be a substitute of mathematics, arts, philosophy or religion.
+The properties of prime numbers, moral and existential questions or the aesthetics of music fail in at least one of those respects and are excluded.
 Only the parts of nature that are accessible through consistent experimental verification can be studied scientifically.
-New experimental techniques extend its reach.
+New experimental techniques have been extending that reach over time.
 
@@
 
-The basic building block in our framework is the idea of a verifiable statement: an assertion that is either true or false for everybody and for which we have an experimental test that will terminate successfully in finite time if and only if the statement is true.
-The basic structures of our general theory need to provide a logic framework to keep track of verifiable statements and their relationships.
+The basic element in our framework is the idea of a verifiable statement: an assertion that is either true or false for everybody and for which we have an experimental test that will terminate successfully in finite time if and only if the statement is true.
+The most basic structures therefore need to provide a logic framework to keep track of what statements are verifiable and their relationships.
 
@@
 It is the most fundamental structure and it is the only one that is axiomatically introduced.
-A scientific theory will consist of a set of statements taken within a logical context with particular logical relationships.
-For example, Newtonian mechanics will use statements like ``the mass of the object is 1 $Kg$'' and ``the acceleration of the object is 1 $m/s^2$''.
-If both of these are true, then the statement ``the force on the object is  1 $N$'' is also true.
-The context defines basic logical relationships (e.g.
+A scientific theory will consist of a set of statements taken within a logical context with particular logical relationships.\footnote{For example, Newtonian mechanics will use statements like ``the mass of the object is 1 $Kg$'' and ``the acceleration of the object is 1 $m/s^2$''.
+If both of these are true, then the statement ``the force on the object is  1 $N$'' is also true.} The context defines basic logical relationships (e.g.
 equivalence, narrowness, compatibility and independence) which form the basis for higher level constructs (e.g.
 ordering, linear and statistical independence).
-Every other structure imposed on the statements will need to ``play'' nice with these fundamental relationships.
+Every other structure imposed on the statements, to be logically consistent, will need to ``play'' nice with these fundamental relationships.
 
@@
 \begin{description}
-	\item not be propositional; the language of the statements, its syntax, its grammar, will be left unspecified; statements will represent the content regardless of how it is expressed
+	\item not be propositional; the language of the statements, its syntax, its grammar, will be left unspecified; statements will represent the content regardless of how it is expressed;
 	\item be algebraic; operations on statements will not form new statements, but describe relationships within a given set of statements; note that most mathematical structures used in science are algebraic (computer science is a notable exception)
-	\item purely formal; it will not not try to capture the semantic of the statements, though the semantic will impose logical relationships that will be formally captured
+	\item purely formal; the semantic of the statements is not captured, though the semantic will impose logical relationships that will be formally captured
 \end{description}
@@
 The existence of this structure is introduced by axioms as they are crossing the lines between physical objects and mathematical ones.
+We currently do not assume there is a single universal context.
 
@@
 A statement is experimentally verifiable if we are provided with a test that, if the statement is true, will always terminate successfully in a finite time.
-The fact that the test has to terminate in a finite time in the positive case and that it may not terminate in the negative case has profound implications: verifiable statements are not closed under the standard logical connectors.
+The fact that the test has to terminate in a finite time in the positive case and that it may not terminate in the negative case has profound implications.
 The negation of a verifiable statement is not in, general, a verifiable statement.
@@
 logical OR) of verifiable statements is a verifiable statement, but not a more then countable infinite one.
-
-\emph{TODO.} The negation of a verifiable statement is not in general a verifiable statement.
-We may be able to verify that ``there exists extra-terrestrial life'' if at some point we find a clear signal.
-Absence of signal, however, will never confirm that ``there is no extra-terrestrial life''.
+Therefore verifiable statements are not closed under the standard logical connectors
 
@@
 
-We will define an experimental domain as a set of verifiable statements that can be expressed as the combination of a countable subset (i.e.
+We define an experimental domain as a set of verifiable statements that can be expressed as the combination of a countable subset (i.e.
 a countable base).
@@
 This will include all statements for which an experimental test is in principle possible, though there is no guarantee of termination.
-A theoretical domain, then, represents all statement to which we can attach a prediction.
+A theoretical domain, then, represents all statements to which we can attach a prediction.
 
 Within a theoretical domain we define the set of possibilities as those statements that, once known to be true, will set the truth value of all other statements.
-Each possibility represents a possible case that is distinguishable experimentally, which corresponds to a unique possible assignment for the experimental and theoretical domain.
-Because of the countable base, the set of possibilities can never be greater than the continuum.
-
-To sum up, an experimental domain consists of a set of statements that can be tested experimentally (e.g.
+Each possibility represents a possible case that is distinguishable experimentally and corresponds to a unique possible assignment for the experimental and theoretical domain.\footnote{For example, an experimental domain consists of a set of statements that can be tested experimentally (e.g.
 ``the animal has whiskers'', ``the mass of the photon is less than $10^{-13} eV$'').
@@
 The possibilities consists of all the possible cases (e.g.
-``the animal is a cat'', ``the mass of the photon is exactly $0.102 10^-{34} eV$ ).
+``the animal is a cat'', ``the mass of the photon is exactly $0.102 10^-{34} eV$ ).} Because of the countable base, the set of possibilities can never be greater than the continuum.
 
@@
 
-An experimental domains provide a natural topology over the possibilities.
-Each verifiable statements can be identified with the set of possibilities compatible with it (e.g.
-the statement ``the mass of the photon is less than $10^{-13} eV$'' is equal to the disjuntion of all statement of the form ``the mass of the photon is exactly x $eV$'' with $x <  10^{-13}$).
-Since verifiable statements are closed under finite conjunction and countable disjunction, the sets corresponding to verifiable statements form a $T_0$ second countable topology.
+An experimental domain provide a natural topology over the possibilities.
+Each verifiable statements can be identified with the set of possibilities compatible with it.\footnote{For example, the statement ``the mass of the photon is less than $10^{-13} eV$'' is equal to the disjuntion of all statement of the form ``the mass of the photon is exactly $x~eV$'' with $x <  10^{-13}$} Since verifiable statements are closed under finite conjunction and countable disjunction, the sets corresponding to verifiable statements form a $T_0$ second countable topology.
 
@@
 Topologies and $\sigma$-algebras provide the foundations for differential geometry, Lie algebras, measure theory, probability theory and many other mathematical tools used in physics and the sciences.
-As we have now have a precise understanding of what they represent, all concepts and proofs in those subject can be understood in terms of experimental verifiability.
+As we now have a precise understanding of what they represent, all concepts and proofs in those subject can be understood in terms of experimental verifiability.
 
@@
 compact sets, all separability axioms, ...).
-It would be interesting to work with people from different fields (i.e.
-logic, foundations of mathematics, philosophy of science, ...) to see whether any aspect of this section would appeal to different communities.
+It would be interesting to reach out to individuals from related fields (i.e.
+logic, foundations of mathematics, philosophy of science, ...) to see whether any aspect of this section would be novel and interesting to their respective communities.
 
-\subsection{Common constructions}
+\subsection{Domain relationships}
 
-With the basic mathematical structure defined, we develop some definitions and constructions that are of general use.
+The study of experimental domains gives us basic constructions that investigate the way different domains can be related or combined.
 
@@
 
-We define two relationships between domain: inference relationships and causal relationships.
+We define two relationships between domains: inference relationships and causal relationships.
 An inference relationship establishes that testing a verifiable statement in one domain is the same as testing a verifiable statement in the other.
-Mathematically it is a map that takes a verifiable statement from the one and returns a verifiable statement from the other that is equivalent to the first.
+Mathematically it is a map that takes a verifiable statement from one and returns a verifiable statement from the other that is equivalent to the first.
 A causal relationship establishes that determining which possibility is true in one domain also determines which possibility is true in the second.
 Mathematically it is a map that given a possibility of the first returns a possibility of the second that is broader than (i.e.
-implied from) the first.
+less specific, true in more cases) the first.
 
-Among the main results we find that there exists an inference implication between two domains if and only if there exists a causal relationship between them.
+One result is that causal relationships must be continuous function in terms of the natural topologies, which justifies why functions in science are always assumed to be ``well-behaved'' (i.e.
+analytically continuous with at most countable discontinuities).
+Another results is that there exists an inference implication between two domains if and only if there exists a causal relationship between them.
 The direction of the inference is the opposite of the causal direction.
 Therefore we say that a domain depends on another if there exists an inference relationship between the first and the second or if there exists a causal relationship between the second and the first.
-Causal relationships must be continuous function in terms of the natural topologies, which justifies why functions in science are always assumed to be ``well-behaved'' (i.e.
-analytically continuous with at most countable discontinuities).
+If the relationships are invertible (i.e.
+each domain depends on the other) then the domains are equivalent.
 
@@
 
+Given two experimental domains, we can create the composite by including all the verifiable statements that can be constructed from them.
+A possibility of the composite domain will determine the truth assignment for all statements of both original domains.
+Depending on the logical relationships between the two domain, we have different cases.
+If the two domains are independent (e.g.
+position along two different directions), then the possibilities are the scalar product of the possibilities.
+If one of the domains depends on the other (e.g.
+the temperature of a mercury column and its height), then the possibilities are the ones of the independent domain.
+If the domains are incompatible (e.g.
+plant species and animal species) then the possibilities are the disjoint union.
+
+%\emph{Product topology naturally emerges}
+
 \subsubsection{Relationship domains}
+
+Given two experimental domains where one depends on the other, we would like to characterize how the relationship is identified experimentally.\footnote{For example, knowing that there is a causal relationship between the temperature and the height of a mercury column, we would like to measure how the two are related.} The main difficulty is that, within a single context, there can only be one such relationship or we would introduce logical inconsistencies.
+Choosing an inference relationship, in fact, means fixing the logical relationship between statements\footnote{For example, ``the temperature is 12 C'' is incompatible with ``the height is 23 mm''} which ultimately means fixing a logical context.
+Identifying the relationship, then, means identifying the right context.
+
+Given a set of contexts, each describing a possible relationship, we can construct an experimental domain for the relationship in which each possibility corresponds to a possible relationship and its context.
+Whether or not the experimental test will exist in practice, the mathematical construction can always be performed resulting in just another domain.
+This means that we can construct relationship domains over relationship domains for any arbitrary higher order relationship, meaning that the mathematical framework is closed.
+
+\emph{Tests for the relationship domain.} The experimental verifiability of relationship domain statements cannot be proven from the verifiability of each domain.
+This formalizes the problem that being able to measure a single point (i.e.
+the temperature and height in a particular circumstance) is not enough to verify the relationship (i.e.
+the temperature and height are always linked in a particular way).
+The ability to experimentally explore the space is what builds confidence, which may not be always possible.
+What constitutes enough evidence is left, as before, outside of the formalism.
 
@@
 
-\subsubsection{Properties and quantities}
+Another basic tool is the ability identify the possibilities of a domain through the values of its properties, since these are the ones that, in practice, we measure.
+Part of the general theory is the formalization of these concepts.
 
-\subsubsection{Ordering theorems for discrete and continuous quantities}
+\subsubsection{Properties and values}
 
-If a scientific theory is a model, then it is a collection of statements with well-defined logical relationships between them.
-For example, a model of the solar systems will be able to say characterized by statements such as ``the earth rotates around
+A property is defined as a map from the possibilities to the set of values that the property can have.
+The map has to be topologically continuous as a measurement on the property should correspond to a measurement over the domain.
+We distinguish between various cases depending on whether the property is defined over all the possibilities and how much is able to distinguish between all the elements
 
-* Formal theories are specializations of more general ones
-* Algebraic structures vs Propositional structures
-* Physical theories as models; the theory of models
-* Difference from ontological theories and subjective theories; accessible information
-* Verifiable statements
-* Meaning of statements can't be captured, requires context
+\subsubsection{Quantities and ordering}
 
-\section{Physics theories}
+Quantities are defined to be ordered properties, values have magnitudes that can be compared.
+The objective is to understand how this ordering is fully defined by the logical relationships between verifiable statement.
+The key insight is that $3 \leq 4$ precisely because ``the quantity is less than 3'' is narrower (i.e.
+more specific) than ``the quantity is less than 4''.
+
+Ordered domains are constructed through references (e.g.
+ticks of a clock, marks of a ruler or levels of a graded recipient) that allow us to tell whether the quantity is before or after itself.
+To have an ordered quantity, the set of references have to satisfy a set of necessary and sufficient conditions that are laid out by three ordering theorems.
+
+Once those conditions are satisfied, the real and integer numbers (i.e.
+continuous and discrete ordering) respectively emerge by requiring either that between two references there is always at least one more or that there are only finitely many.
+
+Of note is that the requirements for ordering itself are quite demanding and are ultimately unlikely to be satisfied for quantities like space and time.
+Which would mean that not only time (or space) would not be chartable by the real line, but that ordering would not be well defined at the finest scale.
+
+\subsubsection{Status and open issues}
+
+Properties and quantities are fairly well developed. The extension to manifolds should be straight forward (require independent quantities to identify each possibility within a ``small'' open set).
+
+\subsection{Further work}
+
+The general theory at this point covers only the topological aspects and needs to be extended to the geometrical/measure theoretical ones.
+The idea should be that diverse aspects, such as geometry, measure theory, probability theory, information theory, can be recovered by characterizing the granularity (i.e.
+accuracy) of the statements.
+As each statement corresponds to a Borel set of possibilities, assigning a size to these sets effectively means giving a size to the statements themselves.
+For geometry, two sets have the same volume because they contain the same number of possible spatial points.
+For probability, borrowing idea from Cox probable inference, we would assign probability only on conditional statements, therefore the $P(s_1 | s_2)$ would measure what fraction of the possibilities compatible with $s_2$ are also compatible with $s_1$.
+For information theory, given two statements the logarithm of the ratio of the size of the possibilities corresponds to the number of yes/no questions that are needed to go from one statement to the other, the relative information.
+
+These ideas still need to be developed in a full fledged theory to see whether they can work all the way through.
+
+\section{Physics core}
+
+This part has not yet been formally developed, though some common ideas are starting to emerge.
+
+\emph{All states are equilibrium states.} In thermodynamics equilibria are simply assumed to exist, without specifying exactly how they are achieved.
+However, in all physical theories, states are assumed to exist, without specifying what are the conditions upon which the system can be characterized by such states.
+The idea that, at the very least, all states must be equilibrium states of processes that happen at a faster scale than the processes than we are considering.\footnote{For example, the position and velocity of a book makes sense on the surface of the earth, at a temperature of 293 Kelvin.
+On the surface of the sun, we would not be able to talk about the position and velocity of a book.} This sets the scale of infinitesimal time intervals.
+
+\emph{States are objective but not absolute.} The same system can be found and studied in different conditions.
+Depending on the circumstances we may be interested in studying different properties of the system.\footnote{Consider a ball.
+If we are studying its motion, the state can be typically be captured by position and momentum of the center of mass.
+If the ball is rotating at high speeds in air, the rotation speed and axis will also need to be part of the state because of the Magnus effect.
+If we are studying the gas within the ball, the state may be given by volume, pressure and temperature.
+If the ball is so tiny and light that random impact of the fluids on the surface will affect the motion, the state will be a statistical ensemble.
+And so on.} Therefore the state will change depending on the situation and our focus, and in this sense it is not absolute.
+However, once the situation and focus are well defined, the state is objective in the sense that everybody will come to the same characterization.
+
+\emph{The tension between isolation and interaction.} Properties of a system can be fully characterized if they are not perturbed by the interaction with the environment, if they are in stable equilibrium.\footnote{For example, the position of the center of mass of a cannonball is not affected by the air and can therefore be characterized by a set of numbers; however, the position of the center of mass of a spek of dust is affected by the air and must be characterized by a probability distribution over the possible values.} To fully characterize a system, then, we would want it to be in perfect isolation.\footnote{Note that full isolation is not possible as this would mean no interaction, not even gravitation, between the system and the environment, to the point that the system has essentially vanished.} However, to be able to measure those quantities we would like to interact with the system.
+To fully access the system, then, we would want it to be in constant interaction with the environment.
+Conceptually, there is a trade-off.
+Specifying the state space means characterizing this trade-off for the specific system in the specific circumstances.
+
+[To be completed]
+
+\section{Physical theories}
 
@@
 We call particle an infinitesimal part of the system, the limit of recursive subdivision.
-The system is then described by a distribution of the state its parts over phase space, which is the only structure that allows coordinate invariant distributions over continuous variables.
+The system is then described by a distribution of the state of its parts over phase space, which is the only structure that allows coordinate invariant distributions over continuous variables.
 
@@
 
-\subsection{Thermodynamics}
+Quantum mechanics follows the same strategy as classical mechanics with the difference that it assumes irreducibility instead of infinitesimal reducibility.
+From this, one can show that the state of the overall system is a complex inner product vector space.
+The deterministic and reversible evolution corresponds to unitary evolution.
+From kinematic equivalence one recovers Lagrangian and potential forces as in the classical case
 
-\subsection{Statistical mechanics}
+[To be completed]
+
+\subsubsection{Quantum state space}
+
+\textbf{Irreducibility.} A system is said irreducible if giving the state of the whole system tells us nothing about the state of its parts.
+For example, given an electron, we can scatter a photon off of it to learn its state.
+Yet, we cannot scatter with only a part of the electron, we only scatter with the whole electron.
+The system is irreducible because we cannot study, and assign a state, to its parts
+
+[To be completed]
+
+% divisible vs reducible vs decomposible
+
+\subsubsection{Status and open issues}
+
+While we consider quantum mechanics to be essentially forward engineered, that is we know the assumptions are necessary and sufficient, there are some conceptual aspects and details that, if clarified, would paint a more satisfactory picture.
+
+[To be completed]
+
+\subsection{Thermodynamics and statistical mechanics}
+
+Reverse engineering work for thermodynamics and statistical mechanics is ongoing.
+
+\subsection{Other theories}
+
+We have not yet attempted to work with field theories, either classical or quantum.
+On the quantum side, there are two obstacles we would need to solve.
+The first is how to characterize spin and the second is the right mathematical tools to use (e.g.
+algebraic geometry?).
+On the classical side, the main conceptual issue is that we start with distributions over $q$ and $p$ so it is not clear what these would represent as distributions over field values.
+
+We haven't yet attempted to work with general relativity as of yet.
+Classical mechanics as derived in this project is already relativistic, in the sense that we derive a metric tensor that is locally Minkowskian.
+However, it is not clear whether that metric has a link to the energy/momentum tensor or it is independent.
 

### 2019-12-15 Articles/Overview/overview.tex
Double version for overview

@@
 
-\title{Overview and Status of the Assumptions of Physics}
+\newif\ifextended
+
+\extendedfalse
+
+\newcommand\fact[2]{\ifextended \emph{#1}.#2 \fi}
+
+\title{Overview and Status of Assumptions of Physics\footnote{http://assumptionsofphysics.org} \\
+\ifextended (Extended version)
+\else (Short version)
+\fi}
 \author{Gabriele Carcassi, Christine A. Aidala}
@@
 
-\emph{Not every subject can be studied by science.} The principle indirectly states that anything that is subjective, contradictory or not evidence based cannot be the subject of scientific investigation.
+\fact{Not every subject can be studied by science} {The principle indirectly states that anything that is subjective, contradictory or not evidence based cannot be the subject of scientific investigation.
 The properties of prime numbers, moral and existential questions or the aesthetics of music fail in at least one of those respects and are excluded.
 Only the parts of nature that are accessible through consistent experimental verification can be studied scientifically.
-New experimental techniques have been extending that reach over time.
+New experimental techniques have been extending that reach over time.}
 
@@
 
-\emph{Note on terminology.} The terminology may depart slightly from what logicians may be accustomed to, simply because our aim is different.
-Namely, our goal is not to study the rule of inference or study what can be formally proven, but simply to keep track of the logical relationships between statements that are assumed to be already given.
+\fact{Note on terminology}{The terminology may depart slightly from what logicians may be accustomed to, simply because our aim is different.
+Namely, our goal is not to study the rule of inference or study what can be formally proven, but simply to keep track of the logical relationships between statements that are assumed to be already given.}
 
-\emph{Statements as primary object.} Every formal theory needs prime objects, elements that will not further be described by the theory.
+\fact{Statements as primary object} {Every formal theory needs prime objects, elements that will not further be described by the theory.
 We will take statements, assertion that are either true or false for everybody, as these objects.
@@
 	\item purely formal; the semantic of the statements is not captured, though the semantic will impose logical relationships that will be formally captured
-\end{description}
+\end{description}}
 
-\emph{Statements cannot exist in isolation.} Statements should, in general, not be regarded as independent of the context they are part of.
+\fact{Statements cannot exist in isolation} {Statements should, in general, not be regarded as independent of the context they are part of.
 Take the statement ``The mass of the electron is 510 $\pm$ 0.5 KeV''.
@@
 If we are performing particle identification in a detector, that statement is assumed to be true.
-The truth values allowed by the same statement depends on context.
+The truth values allowed by the same statement depends on context.}
 
-\emph{Logical context as the fundamental structure.} A logical context is a set of statements with well defined logical relationships.
+\fact{Logical context as the fundamental structure} {A logical context is a set of statements with well defined logical relationships.
 It defines which truth assignments are possible (i.e.
@@
 The existence of this structure is introduced by axioms as they are crossing the lines between physical objects and mathematical ones.
-We currently do not assume there is a single universal context.
+We currently do not assume there is a single universal context.}
 
-\emph{Tautologies, contradictions and contingent statements.} We define tautologies as those statements that are true in every assignment, contradictions as those statements that are false in every assignment and contingent all the remaining.
+\fact{Tautologies, contradictions and contingent statements} {We define tautologies as those statements that are true in every assignment, contradictions as those statements that are false in every assignment and contingent all the remaining.}
 
-\emph{Statement equivalence and Boolean algebra.} Within our formal system, we can distinguish between:
+\fact{Statement equivalence and Boolean algebra} {Within our formal system, we can distinguish between:
 \begin{description}
@@
 \end{description}
-A logical context is a complete Boolean algebra under statement equivalence.
+A logical context is a complete Boolean algebra under statement equivalence.}
 
-\emph{Narrowness and ordering.} A statement is narrower then another if whenever it is true then the other is also true.
+\fact{Narrowness and ordering} {A statement is narrower then another if whenever it is true then the other is also true.
 Narrowness can often be thought as implication, except in a few corner case (e.g.
@@
 Narrowness imposes a partial order to the structure.
-This is the same order associated to the fact that every Boolean algebra is a complemented distributed lattice.
+This is the same order associated to the fact that every Boolean algebra is a complemented distributed lattice.}
 
-\emph{Compatibility and independence.} Two statements are said to be compatible if they can both be true in the same assignment.
+\fact{Compatibility and independence} {Two statements are said to be compatible if they can both be true in the same assignment.
 Two statements are independent if fixing the truth value of one does not restrict the truth value of the other.
-Independence, like linear independence or statistical independence of which it is the foundation, is a property of a set of statements and is not transitive.
+Independence, like linear independence or statistical independence of which it is the foundation, is a property of a set of statements and is not transitive.}
 
@@
 
-\subsubsection{Experimental domains}
+[To be completed]
+
+\subsubsection{Experimental and theoretical domains}
 
@@
 
+[To be completed]
+
 \subsubsection{Topologies and $\sigma$-algebras}
@@
 As we now have a precise understanding of what they represent, all concepts and proofs in those subject can be understood in terms of experimental verifiability.
+
+[To be completed]
 
@@
 
+[To be completed]
+
 \subsubsection{Composite domains}
@@
 
-\emph{Tests for the relationship domain.} The experimental verifiability of relationship domain statements cannot be proven from the verifiability of each domain.
+\fact{Tests for the relationship domain} {The experimental verifiability of relationship domain statements cannot be proven from the verifiability of each domain.
 This formalizes the problem that being able to measure a single point (i.e.
@@
 The ability to experimentally explore the space is what builds confidence, which may not be always possible.
-What constitutes enough evidence is left, as before, outside of the formalism.
+What constitutes enough evidence is left, as before, outside of the formalism.}
 
@@
 This part has not yet been formally developed, though some common ideas are starting to emerge.
+The main idea comes from the realization that, when defining a system, the notions of states, processes, time, evolution laws, interaction with the environment and equilibria are all interrelated and can't be given independently.
+Roughly speaking, to define a system we have to choose a boundary that separates it from the environment.
+Only the quantities that are unaffected by the interaction with the environment can, in those circumstances, be thought as properties of the system.
+This in turns is linked to what we can define as state for the system, which is linked to the laws we can write.
+The processes and the laws are naturally linked to what is assumed about time.
 
-\emph{All states are equilibrium states.} In thermodynamics equilibria are simply assumed to exist, without specifying exactly how they are achieved.
+The idea, then, is that giving a set of states is actually giving a much richer structure.
+It means defining a set of interaction at the boundary, it means defining equilibria with such interactions, it means defining time-scales below which and above which our description fails or is not interesting, it means defining how probing the system may affect those assumptions.\footnote{If we are studying a molecule, we may impose boundaries and processes such that that molecule and its constituent remain stable, which means, for example, interaction energies lower than the chemical bonds.} The plan is, at some point, to formalize these objects and relationships to provide a basic mathematical structure that all physical theories must adhere to.
+
+\fact{All states are equilibrium states} {In thermodynamics equilibria are simply assumed to exist, without specifying exactly how they are achieved.
 However, in all physical theories, states are assumed to exist, without specifying what are the conditions upon which the system can be characterized by such states.
 The idea that, at the very least, all states must be equilibrium states of processes that happen at a faster scale than the processes than we are considering.\footnote{For example, the position and velocity of a book makes sense on the surface of the earth, at a temperature of 293 Kelvin.
-On the surface of the sun, we would not be able to talk about the position and velocity of a book.} This sets the scale of infinitesimal time intervals.
+On the surface of the sun, we would not be able to talk about the position and velocity of a book.} This sets the scale of infinitesimal time intervals.}
 
-\emph{States are objective but not absolute.} The same system can be found and studied in different conditions.
+\fact{States are objective but not absolute} {The same system can be found and studied in different conditions.
 Depending on the circumstances we may be interested in studying different properties of the system.\footnote{Consider a ball.
@@
 And so on.} Therefore the state will change depending on the situation and our focus, and in this sense it is not absolute.
-However, once the situation and focus are well defined, the state is objective in the sense that everybody will come to the same characterization.
+However, once the situation and focus are well defined, the state is objective in the sense that everybody will come to the same characterization.}
 
-\emph{The tension between isolation and interaction.} Properties of a system can be fully characterized if they are not perturbed by the interaction with the environment, if they are in stable equilibrium.\footnote{For example, the position of the center of mass of a cannonball is not affected by the air and can therefore be characterized by a set of numbers; however, the position of the center of mass of a spek of dust is affected by the air and must be characterized by a probability distribution over the possible values.} To fully characterize a system, then, we would want it to be in perfect isolation.\footnote{Note that full isolation is not possible as this would mean no interaction, not even gravitation, between the system and the environment, to the point that the system has essentially vanished.} However, to be able to measure those quantities we would like to interact with the system.
+\fact{The tension between isolation and interaction.} {Properties of a system can be fully characterized if they are not perturbed by the interaction with the environment, if they are in stable equilibrium.\footnote{For example, the position of the center of mass of a cannonball is not affected by the air and can therefore be characterized by a set of numbers; however, the position of the center of mass of a spek of dust is affected by the air and must be characterized by a probability distribution over the possible values.} To fully characterize a system, then, we would want it to be in perfect isolation.\footnote{Note that full isolation is not possible as this would mean no interaction, not even gravitation, between the system and the environment, to the point that the system has essentially vanished.} However, to be able to measure those quantities we would like to interact with the system.
 To fully access the system, then, we would want it to be in constant interaction with the environment.
 Conceptually, there is a trade-off.
-Specifying the state space means characterizing this trade-off for the specific system in the specific circumstances.
+Specifying the state space means characterizing this trade-off for the specific system in the specific circumstances.}
 
@@
 
-\emph{Composite states are distributions over particle states.} Under infinitesimal reducibility, if $\mathcal{C}$ is the state space of the while system and $\mathcal{S}$ is the state space of the particles, then each state $c \in \mathcal{C}$ for the full system is identified by a distribution $\rho : \mathcal{S} \to \mathbb{R}$ over the states of infinitesimal parts.
+\fact{Composite states are distributions over particle states} {Under infinitesimal reducibility, if $\mathcal{C}$ is the state space of the while system and $\mathcal{S}$ is the state space of the particles, then each state $c \in \mathcal{C}$ for the full system is identified by a distribution $\rho : \mathcal{S} \to \mathbb{R}$ over the states of infinitesimal parts.
 The function is real to signify that we can associate arbitrarily small amounts to each particle states.
-The distribution tells us how many particles can be found in each particle state.
+The distribution tells us how many particles can be found in each particle state.}
 
-\emph{Continuous time implies particle state space is a manifold.} If time is assumed to be continuous, then the state space $\mathcal{S}$ is a manifold.
+\fact{Continuous time implies particle state space is a manifold} {If time is assumed to be continuous, then the state space $\mathcal{S}$ is a manifold.
 That is, each particle state $s \in U \subseteq \mathcal{S}$ can be identified by a set of continuous quantities $\xi^a : U \to \mathbb{R}$ which we call state variables.
 It is a consequence of the general theory that, if we wanted to write a trajectory $\xi^a(t)$, then this must be a topologically continuous function.
-Therefore, conceptually, once we assume time to be a real number, the state space of the particles must be charted by a set of real quantities: the topology that is given to time reasserts itself on all the quantities that will depend on time.
+Therefore, conceptually, once we assume time to be a real number, the state space of the particles must be charted by a set of real quantities: the topology that is given to time reasserts itself on all the quantities that will depend on time.}
 
-\emph{Densities imply differentiability.} As we can write the state $s(\xi^a)$ as a function of the state variables $\xi^a$, we will also want to write the density $\rho(s(\xi^a)) = \rho(\xi^a)$ as a function of the state variables.
+\fact{Densities imply differentiability} {As we can write the state $s(\xi^a)$ as a function of the state variables $\xi^a$, we will also want to write the density $\rho(s(\xi^a)) = \rho(\xi^a)$ as a function of the state variables.
 This requires the state variables to be differentiable with respect to the distribution and with each other.
@@
 Note that any time evolution that would map a density to another density will not only have to be continuous, but also differentiable.
-In short: assuming density distributions as our primary objects, instead of points, justifies why differentiable structures and differentiable time evolution is present in physics.
+In short: assuming density distributions as our primary objects, instead of points, justifies why differentiable structures and differentiable time evolution is present in physics.}
 
-\emph{Coordinate invariant densities imply phase-space.} As particles states must be defined independently of coordinate systems, they must be invariant under coordinate transformations.
+\fact{Coordinate invariant densities imply phase-space} {As particles states must be defined independently of coordinate systems, they must be invariant under coordinate transformations.
 That is, $s(\xi^a) = s(\hat{\xi}^b(\xi^a))$.
@@
 Mathematically, symplectic manifolds are the only spaces that allow us to define coordinate invariant densities.
-We can introduce a constant $\hbar$ such that $p_i = \hbar k_i$ which at this point is simply an arbitrary choice of unit.
+We can introduce a constant $\hbar$ such that $p_i = \hbar k_i$ which at this point is simply an arbitrary choice of unit.}
 
-These results gives us a precise understanding of what each layer of the mathematical structure corresponds to physically:
+\fact{Precise physics-math correspondence} {These results gives us a precise understanding of what each layer of the mathematical structure corresponds to physically:
 \begin{description}[noitemsep]
@@
 	\item \emph{Symplectic manifold:} we can define a distribution over the states that are coordinate independent.
-\end{description}
+\end{description}}
 
@@
 
-\emph{Reversibility as reconstructing the past.} This definition of reversibility should not be confused with time symmetry or the ability to undo the change.
+\fact{Reversibility as reconstructing the past} {This definition of reversibility should not be confused with time symmetry or the ability to undo the change.
 The first one means that if one substitute $t \to -t$ the dynamics is invariant (i.e.
@@
 It is mathematically more elegant as it is the dual of determinism.
-It connects directly with information theory, as it states that the present has enough information to reconstruct the past, which gives a logical bridge to connect reversibility with information entropy and thermodynamics/statistical mechanics entropy.
+It connects directly with information theory, as it states that the present has enough information to reconstruct the past, which gives a logical bridge to connect reversibility with information entropy and thermodynamics/statistical mechanics entropy.}
 
-\emph{Determinism and reversibility imply Hamiltonian mechanics.} In the context of infinitesimal reducibility, determinism and reversibility means that all the particles and only the particle within a given initial state $\rho(s(t)) = \rho(s(t+\Delta t))$ will be mapped to a particular final state.
+\fact{Determinism and reversibility imply Hamiltonian mechanics} {In the context of infinitesimal reducibility, determinism and reversibility means that all the particles and only the particle within a given initial state $\rho(s(t)) = \rho(s(t+\Delta t))$ will be mapped to a particular final state.
 That is, the densities will be preserved over time.
@@
 As time is continuous, we can find a function $H : \mathcal{S} \to \mathbb{R}$ which acts as a time generator, which we recognize as being the Hamiltonian.
-This is essentially Louisville's theorem applied in reverse.
+This is essentially Louisville's theorem applied in reverse.}
 
-\emph{Invariant densities under time transformations imply extended phase-space.} So far, time has been treated as a parameter.
+\fact{Invariant densities under time transformations imply extended phase-space} {So far, time has been treated as a parameter.
 If we introduce coordinate transformation that mix time, however, this is no longer possible.
@@
 In other words, there is no way to have time transformations in Hamiltonian mechanics without requiring a four-momentum co-vector.
-Special relativity is not an independent choice: it is already partially baked into the two assumptions we took.
+Special relativity is not an independent choice: it is already partially baked into the two assumptions we took.}
 
-\emph{Extended phase-space imply classical anti-particles.} The trajectories in the extended phase space are in terms of an affine parameter $s$ and not time: $q^\alpha(s)$.
+\fact{Extended phase-space imply classical anti-particles} {The trajectories in the extended phase space are in terms of an affine parameter $s$ and not time: $q^\alpha(s)$.
 As the motion is still deterministic and reversible, though, time $t(s)$ must be monotonic with respect to the affine parameter.
@@
 Because the relationship is monotonic, aligned states can only connect to other aligned states and anti-aligned states can only connect to other anti-aligned states.
-We call aligned states the particle states and the anti-aligned states the anti-particle states.
+We call aligned states the particle states and the anti-aligned states the anti-particle states.}
 
-\emph{Time evolution is better expressed by the Hamiltonian constraint.} The temporal degree of freedom introduces two variables but, since one is energy, there is an additional constraint since the value of the energy is linked to the other variables through the Hamiltonian.
+\fact{Time evolution is better expressed by the Hamiltonian constraint} {The temporal degree of freedom introduces two variables but, since one is energy, there is an additional constraint since the value of the energy is linked to the other variables through the Hamiltonian.
 This link is better expressed by the Hamiltonian constraint $\mathcal{H} : \mathcal{\bar{S}} \to \mathbb{R}$ which is a function over the extended phase space, just like the Hamiltonian was a function over the standard phase space.
@@
 The Hamiltonian constraint, unlike the standard Hamiltonian, is invariant under time transformation.
-It is the generator of the evolution in terms of an affine parameter.
+It is the generator of the evolution in terms of an affine parameter.}
 
-\emph{Hamiltonian constraint as operator on distributions.} If $\rho$ is the distribution for the whole system, then $\mathcal{H} \rho = 0$.
+\fact{Hamiltonian constraint as operator on distributions} {If $\rho$ is the distribution for the whole system, then $\mathcal{H} \rho = 0$.
 In fact, the distribution can only be non-zero on the constrain, for which the Hamiltonian constraint is zero.
 So, for each point in the extended phase space, either the distribution is zero or the Hamiltonian constrain is zero.
-Note that this is the same form one has in quantum mechanics for the Klein-Gordon and Dirac equations $\mathcal{H} |\psi\rangle = 0$, which in fact can be understood as Hamiltonian constraints that fix the norm of the momentum four-vector and four-velocity.
+Note that this is the same form one has in quantum mechanics for the Klein-Gordon and Dirac equations $\mathcal{H} |\psi\rangle = 0$, which in fact can be understood as Hamiltonian constraints that fix the norm of the momentum four-vector and four-velocity.}
 
-\emph{Determinism and reversibility is equivalent to conservation of information entropy.} Given a distribution, its information entropy is given by $I = - \int \rho \log \rho \, d\xi^n$.
+\fact{Determinism and reversibility is equivalent to conservation of information entropy} {Given a distribution, its information entropy is given by $I = - \int \rho \log \rho \, d\xi^n$.
 In general, information entropy is not invariant under change of variable: $- \int \rho(\hat{\xi}^b) \log \rho(\hat{\xi}^b) \, d\hat{\xi}^n = - \int \rho(\xi^a) \log \rho(\xi^a) \, d\xi^n - \int \rho(\xi^i) \log \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a}\right| \, d\xi^n$.
@@
 This should be intuitive: the amount of information needed to identify an element within the initial or the final distribution is the same if the evolution is deterministic and reversible.
-Deterministic and reversible evolution, Hamiltonian evolution and conservation of information entropy, then, are one and the same.
+Deterministic and reversible evolution, Hamiltonian evolution and conservation of information entropy, then, are one and the same.}
 
-\emph{Conservation of information entropy leads to classical uncertainty principle.} As information entropy is conserved, a Hamiltonian process can only map a distribution to one that has the same information entropy.
+\fact{Conservation of information entropy leads to classical uncertainty principle} {As information entropy is conserved, a Hamiltonian process can only map a distribution to one that has the same information entropy.
 If we consider a single degree of freedom $(q, p)$ and fix the information entropy $I_0$, then the distribution that minimizes the product of the standard deviation is the Gaussian.
@@
 During the evolution, we have:
-$$ \sigma_q \sigma_p \geq \frac{e^{I_0}}{2\pi e} $$.
+$$ \sigma_q \sigma_p \geq \frac{e^{I_0}}{2\pi e} $$.}
 
@@
 
-Note that we use $(q^\alpha, p_\alpha)$ for position and momentum and $(x^\alpha, u^\alpha)$ for position and velocity even though $q^\alpha = x^\alpha$.
+\fact{Note on notation} {We use $(q^\alpha, p_\alpha)$ for position and momentum and $(x^\alpha, u^\alpha)$ for position and velocity even though $q^\alpha = x^\alpha$.
 We do this because $\frac{\partial}{\partial q^\alpha} \neq \frac{\partial}{\partial x^\alpha}$ since one is taken at constant $p_\alpha$ and the other at constant $u^\alpha$.
-This is a tremendous source of confusion which is avoided by the use of the two symbols.
+This is a tremendous source of confusion which is avoided by the use of the two symbols.}
 
-\emph{Kinematic equivalence means having an invertible relationship between velocity and momentum.} If the state of the system $(q^\alpha, p_\alpha)$ is enough to determine the trajectory $x^\alpha(s)$, then it will be enough to determine the position and velocity $(x^\alpha, u^\alpha)$.
+\fact{Kinematic equivalence means having an invertible relationship between velocity and momentum} {If the state of the system $(q^\alpha, p_\alpha)$ is enough to determine the trajectory $x^\alpha(s)$, then it will be enough to determine the position and velocity $(x^\alpha, u^\alpha)$.
 These will be enough to reconstruct the state as they span a space of equal dimension.
-As $q^\alpha = x^\alpha$, then $u^\alpha = u^\alpha(q^\beta, p_\gamma)$ is invertible.
+As $q^\alpha = x^\alpha$, then $u^\alpha = u^\alpha(q^\beta, p_\gamma)$ is invertible.}
 
-\emph{Kinematic equivalence leads to Lagrangian mechanics.} Since $u^\alpha(q^\beta, p_\gamma)$, we can write $\mathcal{L} = \dot{q}^\alpha p_\alpha - \mathcal{H}$ as a function of $(x^\alpha, u^\alpha)$.
+\fact{Kinematic equivalence leads to Lagrangian mechanics} {Since $u^\alpha(q^\beta, p_\gamma)$, we can write $\mathcal{L} = \dot{q}^\alpha p_\alpha - \mathcal{H}$ as a function of $(x^\alpha, u^\alpha)$.
 That is the Lagrangian of the system.
@@
 Using Hamilton's equations, we have $u^i = \frac{dx^i}{dt} = c \frac{p}{|p|}$ which is not invertible.
-While we can write the Lagrangian, this cannot be expressed in terms of position and velocity.
+While we can write the Lagrangian, this cannot be expressed in terms of position and velocity.}
 
-\emph{Kinematic equivalence leads to linear relationship between velocity and momentum.} Being able to go from $(q^\alpha, p_\alpha)$ to $(x^\alpha, u^\alpha)$ and vice-versa is not enough.
+\fact{Kinematic equivalence leads to linear relationship between velocity and momentum} {Being able to go from $(q^\alpha, p_\alpha)$ to $(x^\alpha, u^\alpha)$ and vice-versa is not enough.
 We need to express the density $\rho$ in terms of position and velocity as well.
 This means expressing area elements $dq^\alpha dp_\alpha$ in terms of $dx^\alpha$ and $du^\alpha$.
-Since $dx^\alpha = dq^\alpha$, we must have $dp_\alpha = m g_{\alpha \beta} du^\beta$ where $g_{\alpha \beta}$ is an arbitrary linear function and $m$ is an arbitrary constant of proportionality.
+Since $dx^\alpha = dq^\alpha$, we must have $dp_\alpha = m g_{\alpha \beta} du^\beta$ where $g_{\alpha \beta}$ is an arbitrary linear function and $m$ is an arbitrary constant of proportionality.}
 
-\emph{Linear relationship between velocity and momentum leads to massive particles under potential forces.} We can integrate the relationship $dp_\alpha = m g_{\alpha \beta} du^\beta$ and find $p_\alpha = m g_{\alpha \beta} u^\beta + A_\alpha(x^\gamma)$ where $A_\alpha$ are arbitrary functions.
+\fact{Linear relationship between velocity and momentum leads to massive particles under potential forces} {We can integrate the relationship $dp_\alpha = m g_{\alpha \beta} du^\beta$ and find $p_\alpha = m g_{\alpha \beta} u^\beta + A_\alpha(x^\gamma)$ where $A_\alpha$ are arbitrary functions.
 Since $u^\alpha = \frac{dq^\alpha}{ds} = \frac{\partial \mathcal{H}}{\partial p_\alpha}$, we can take the previous relationship, integrate it again and find $\mathcal{H} = \frac{1}{2m} (p_{\alpha} - A_\alpha) g^{\alpha \beta} (p_{\beta} - A_\beta) + V$.
-We recognize this to be the Hamiltonian for a massive particle under potential forces.
+We recognize this to be the Hamiltonian for a massive particle under potential forces.}
 
-\emph{Inertial mass measures the number of states per unit of velocity.} Inertial mass is usually introduced using the Newtonian relationship $F=ma$: the lower the mass, the easier is accelerate the body.
+\fact{Inertial mass measures the number of states per unit of velocity} {Inertial mass is usually introduced using the Newtonian relationship $F=ma$: the lower the mass, the easier is accelerate the body.
 If the mass where zero, then, we would expect the body to be infinitely easy to accelerate.
@@
 If there are more states per unit of velocity, then the body is more difficult to accelerate as we have to go through more states, which recover the Newtonian intuition.
-However, if the mass is zero, set of states defined in a unit of velocity has measure zero: we cannot change state, we cannot accelerate.
+However, if the mass is zero, set of states defined in a unit of velocity has measure zero: we cannot change state, we cannot accelerate.}
 
-\emph{The speed of light is a ration between the possible states in a unit of space over the possible states in a unit of time.} Regarding the constant $c$ as a speed is problematic.
+\fact{The speed of light is a ration between the possible states in a unit of space over the possible states in a unit of time} {Regarding the constant $c$ as a speed is problematic.
 It is unclear why a speed, which is a property of motion, would be part of the metric of the space-time, which is defined regardless of motion and of what is there in the space itself.
@@
 Therefore, over a finite time, we must have $\frac{\Delta x}{\Delta t} \leq c$.
-Thinking about the constant $c$ as relating the count of states in space over time is more fundamental.
+Thinking about the constant $c$ as relating the count of states in space over time is more fundamental.}
 
-\emph{The Plank constant, the speed of light and the inertial mass serve to transfer the measure of states from one quantity to another.} As we saw initially, $(q, k)$ are the quantities on which we can measure the number of states.
+\fact{The Planck constant, the speed of light and the inertial mass serve to transfer the measure of states from one quantity to another} {As we saw initially, $(q, k)$ are the quantities on which we can measure the number of states.
 The constant $\hbar$ serves to transfer the count from $dk$ to $dp$.
@@
 The speed of light served to transfer the count from $dt$ to $dx$.
-They are book-keeping tools we use as we express states over different quantities which allow us to express in different ways the fundamental relationship coming from determinism and reversibility: the count of states must remain the same.
+They are book-keeping tools we use as we express states over different quantities which allow us to express in different ways the fundamental relationship coming from determinism and reversibility: the count of states must remain the same.}
 
@@
 
-\emph{Physical meaning of the Poisson brackets.} We still miss an understanding of the Poisson brackets as a generalization of Jacobian determinants.
+\fact{Physical meaning of the Poisson brackets} {We still miss an understanding of the Poisson brackets as a generalization of Jacobian determinants.
 In principle, they should tells us how densities over a degree of freedom formed by two variables transform.
-We haven't found the right approach.
+We haven't found the right approach.}
 
-\emph{Densities as measure theoretic derivatives.} As we want finite objects to be our starting points, densities should be the ration of amount of ``stuff'' divided by the size of the region.
+\fact{Densities as measure theoretic derivatives} {As we want finite objects to be our starting points, densities should be the ration of amount of ``stuff'' divided by the size of the region.
 This, in  measure theory, is known as the Radon-Nikodym derivative.
 We want this idea to be combined with statistics and differential geometry, so that we can talk about marginal distributions.
-It seems that gemoetric measure theory at least partially addresses that, which would be the first step for a full formalization of this section of the project.
+It seems that gemoetric measure theory at least partially addresses that, which would be the first step for a full formalization of this section of the project.}
 
-\emph{Relationship between metric tensor and symplectic form.} The way the derivation works, the metric tensor is a secondary object that should be understood as being applied to a differential of position and a differential of velocity (e.g.
+\fact{Relationship between metric tensor and symplectic form} {The way the derivation works, the metric tensor is a secondary object that should be understood as being applied to a differential of position and a differential of velocity (e.g.
 $dx^\alpha g_{\alpha \beta} du^\beta$).
 This gives us the number of states on an area defined by the differentials.
-It is unclear whether the curvature of $g_{\alpha \beta}$ is linked to the geometry set by the symplectic form.
+It is unclear whether the curvature of $g_{\alpha \beta}$ is linked to the geometry set by the symplectic form.}
 
@@
 Yet, we cannot scatter with only a part of the electron, we only scatter with the whole electron.
-The system is irreducible because we cannot study, and assign a state, to its parts
+The system is irreducible because we cannot study, and assign a state, to its parts.
+
+\fact{Divisibility vs reducibility vs decomposability} {There are three concepts that are often confused to one another.
+Divisibility is the possibility to divide a system into two independent ones.
+That is, we have a time evolution such that we start from a state of a system and we end up with two states of independent systems.
+Reducibility is the ability to describe a system as the composition of more systems.
+That is, giving the state of the system at one time is the same as giving the state of the parts at the same time.
+These are independent properties.\footnote{For example, the Planarian worm is divisible into two worms but is not reducible to two worms.
+A muon is divisible (i.e.
+it decays) into an electron and two neutrinos but is not reducible to them.
+A magnetic is reducible to a north and a south pole but is not divisible into them.
+A proton is reducible to quarks and gluons but is not divisible into them.} Decomposability is the ability to treat one object as the composition of others.
+We take two objects or descriptions and combine them formally to give a new one.
+This is again independent from the rest and often is just mathematical convenience.\footnote{For example, velocity in any direction can be decomposed into the components along the axis, but is not divisible or reducible to those components.
+A pure state is decomposable into a linear combination of a basis but is not reducible or divisible into the basis.}}
 
@@
 On the quantum side, there are two obstacles we would need to solve.
-The first is how to characterize spin and the second is the right mathematical tools to use (e.g.
+The first is how to characterize relativistic spin and the second is the right mathematical tools to use (e.g.
 algebraic geometry?).

### 2019-12-26 Proceedings/2019-Quantum2019/CarcassiQuantum2019Proceedings.tex
Responding to reviewer's comments

@@
 The internal motion is then characterized by random variables, which still need to allow for the definition of an invariant density, and therefore they must, for a degree of freedom, form a pair $(U,V)$.
-As the value of the variable is completely unknown, the distribution will be uniform.
-We can fix the density $\rho=1$ to be unitary and the standard deviations $\sigma_U=\sigma_V$ to be equal as this internal relabelling cannot affect the state of the system as a whole.
+As the value of the variable is completely unknown\footnote{One may be tempted to distinguish between the following cases: the internal dynamics is unknown, the internal dynamics is unknowable, the internal dynamics is absent/not defined.
+In fact, different interpretations of quantum mechanics may take different positions.
+However, one has to be clear how those cases are going to differ experimentally, or this distinction remains within the philosophical realm.
+In our case, we are limiting ourselves to posit that the processes we are studying are not sensitive to a further description of the system, which makes a possible internal dynamics automatically unknown.
+It could be that we will find new processes in the future, it could be that such processes are forever outside of what we can control experimentally or it could be that they do not even exist.
+We cannot know which case we are in: all we is that know we are not sensitive.
+As we do not have a way to tell, these cases are not distinguishable to us experimentally and they can be treated as the same case.}, the distribution will be uniform.\footnote{The addition of extra degrees of freedom that do not alter the problem but make it easier to study is something common in math and physics.
+It is common in probability theory, where adding random variables is a standard technique in proofs.
+It is common in analysis when extending a real integral to the complex plane.
+It is common in statistical mechanics in order to study system/environment interactions.
+It is common in quantum mechanics with purification.
+In all these cases, we are extending our space while making sure that the result does not depend on the particulars of how it was extended.
+Adding an internal dynamics to which we are insensitive is just another instance.} We can fix the density $\rho=1$ to be unitary and the standard deviations $\sigma_U=\sigma_V$ to be equal as this internal relabelling cannot affect the state of the system as a whole.
 The transformations that will affect the system as a whole are the changes in variance and correlation as these would change the value of the integral $\int \rho dudv \propto \sigma_U\sigma_V = \sigma_U^2$.
@@
 Therefore, when comparing classical and quantum mechanics, we should be comparing the properties of our wave function $\psi(q^i)$ to the matter distribution $\rho(q^i, k_i)$, not to points of phase space.
+Comparing the two mechanics, in fact, means comparing how the same system would be different under the different assumptions.
+So we have to compare finite systems to finite systems.
+In both cases, the ``particle'' is the smallest part of the system to which we can assign a state.
+Yet, in quantum mechanics it is itself the indivisible system, while a classical particle is not a system: it is the limit of recursive subdivision.
+While this different comparison may at first seem awkward, as it is not the usual framing, it leads to a more stringent parallel between the mathematical elements of classical and quantum mechanics and, in turn, to a better intuition.
+
+For example, it makes it clear that the position of classical objects (e.g.
+planets, cannonballs, drops of water and even ourselves) cannot be fully described by a single number, as it is often said.
+All these objects have extent both in space and momentum.
 A point-particle is the limit where the extent of the distribution can be considered small, and therefore the center of mass is enough for our description.
 This limit can be performed within both classical and quantum mechanics if the scale of the problem allows, and, because of Ehrenfest's theorem, the behavior is the same.
-In light of the assumptions, it should be obvious this should be the case: the difference between classical and quantum is in the description of the parts; if the length scales are much greater than those of the system, then it doesn't matter what we assume about the parts.
+In light of the assumptions, it should be obvious this should be the case: the difference between classical and quantum is in the description of the parts; if the length scales of interest are much greater than those of the system, then it does not matter what we assume about the parts.
 

### 2019-12-28 Articles/Overview/overview.tex
Preparing for first draft release

@@
 
-\newcommand\fact[2]{\ifextended \emph{#1}.#2 \fi}
+\newcommand\fact[2]{\ifextended \emph{#1}. #2 \fi}
 
@@
 It will contain a summary section at the beginning and a titled paragraph for each of the main idea or finding.
-
-\subsection{Index}
-To this paper, not to the work.
-
-- Goals and method
-Goals: Understanding, have a better sense of the math, give a background for new theories...
-Methods: reverse engineering vs forward engineering, holistic
-
-- General theory
-relationship between math and physics
-logic of verifiable statements
-properties and quantities
-accuracy
-differentiability
-
-- Fundamental physics
-states and processes
-determinism/reversibility: isomorphism in the category
-entropy
-
-- Physical theories
-Classical mechanics
-Quantum mechanics
-Thermodynamics
-Statistical mechanics
 
@@
 As each topic is investigated by the project, it generally goes through three phases.
-We mention them here as they will give context to the status of the project and to people who may want to contribute.
+Reverse engineering, where we try to find what assumptions are beneath a theory.
+Forward engineering, where we try to check whether the theory can be recovered from the assumptions we found.
+Formalization, where we try to codify in mathematical language the assumptions and the derivation.
 
-\textbf{Reverse engineering.} The goal of this phase is to deconstruct an already existing theory or mathematical structure to understand what each piece is supposed to represent physically.
-One way to conduct such investigation is to ask what would happen if a particular characteristic of the framework would fail.\footnote{For example, a line of reasoning could be as follows: Hamilton's equations are differential; if they weren't the mapping of densities and areas would not be well defined; maybe Hamiltonian mechanics is not about mapping points but about mapping densities or areas.} This usually leads to a set of necessary assumptions about the physical system one is describing, the ones one must take for the particular description to apply.
+\fact{Reverse engineering} {The goal of this phase is to deconstruct an already existing theory or mathematical structure to understand what each piece is supposed to represent physically.
+One way to conduct such investigation is to ask what would happen if a particular characteristic of the framework would fail.\footnote{For example, a line of reasoning could be as follows: Hamilton's equations are differential; if they weren't the mapping of densities and areas would not be well defined; maybe Hamiltonian mechanics is not about mapping points but about mapping densities or areas.} This usually leads to a set of necessary assumptions about the physical system one is describing, the ones one must take for the particular description to apply.}
 
-\textbf{Forward engineering.} The goal of this phase is to construct an argument from the necessary assumptions outlined in the previous phase to rederive the framework at hand.
+\fact{Forward engineering} {The goal of this phase is to construct an argument from the necessary assumptions outlined in the previous phase to rederive the framework at hand.
 This checks whether the necessary assumptions are also sufficient \footnote{For example, if we started simply by assuming we have density distributions mapped in time, this would give us all possible differentiable evolutions, not necessarily Hamiltonian ones.}.
-It ends when a set of necessary and sufficient conditions to rederive the theory are found, meaning that the full theory has been physically characterized.
+It ends when a set of necessary and sufficient conditions to rederive the theory are found, meaning that the full theory has been physically characterized.}
 
-\textbf{Formalization.} The goal of this phase is to take the previous set of arguments and formalize them in pure mathematical language.
+\fact{Formalization} {The goal of this phase is to take the previous set of arguments and formalize them in pure mathematical language.
 This forces us to clarify all the starting points and what parts of the arguments are truly formal and what are physical.
-Unfortunately, this can ultimately be done only by starting from the very foundations\footnote{For example, before formalizing Hamiltonian mechanics we have to formalize differentiable manifolds, real numbers and topological spaces.} though we are at a point where an overall tentative structure is emerging and tentative formalization can be attempted on higher structures.
+Unfortunately, this can ultimately be done only by starting from the very foundations\footnote{For example, before formalizing Hamiltonian mechanics we have to formalize differentiable manifolds, real numbers and topological spaces.} though we are at a point where an overall tentative structure is emerging and tentative formalization can be attempted on higher structures.}
 
@@
 
-[To be completed]
+\fact{Different algebra}{Verifiable statements do not follow the standard Boolean algebra.
+The negation of a verifiable statement is not necessarily verifiable since the test is not guaranteed to terminate.
+The conjunction of a finite set of verifiable statements is necessarily verifiable since we can run the tests one by one and, if all are successful, the conjunction is also verified.
+This cannot be extended to an infinite set since the test would not terminate.
+The disjunction of a countable set of verifiable statements is necessarily verifiable since, once one test terminates, the disjunction is verified and we can stop.
+This cannot be extended to an infinite set that is more than countable since we would not be able to find the test that terminates in a finite time.}
+
+\fact{Decidable statements}{These are statements that can be verified and falsified.
+That is, their test always terminates.
+These can be defined from verifiable statements and they follow a finite Boolean algebra: the negation, finite conjunction and finite disjunction of decidable statements is also decidable.}
 
@@
 
-[To be completed]
+\fact{Maximum cardinality for the possibilities}{Because the domain is generated by a countable set of verifiable statements, the possibilities must be distinguished with a countable sequence of boolean values.
+This means that the cardinality of the possibilities cannot exceed that of the continuum.
+This means that higher order infinities and all associated problems do not play a role in science.}
 
@@
 
-[To be completed]
+\fact{Interior, exterior and boundary}{Given a test associated to a theoretical statement, it will either terminate successfully, terminate unsuccessfully or never terminate.
+Each Borel set is associated with a theoretical statement, which means it is associated with a test.
+The interior of the Borel set represents the cases in which the test will terminate successfully; the exterior represents the cases in which the tests will terminate unsuccessfully; the boundary represents the cases in which the test will not terminate.}
+
+\fact{Decidable and undecidable statements}{A theoretical statement is decidable if the associated test always terminates, which means the boundary is empty.
+This corresponds to a clopen set.
+It is undecidable if the associated test never terminates, which means the boundary is the full set.\footnote{For example, ``there are three apples in the bowl'' is a decidable statement, while ``the mass of the electron is a rational number when expressed in eV'' is undecidable.}}
 
@@
 
-[To be completed]
+\fact{Continuity and inference}{An inference relationship establishes that testing a verifiable statement in one domain tells whether a statement in the other is true or false.
+As these statements are equivalent, the experimental test on one constitutes an experimental test on the other, which means the second statement is verifiable.
+One cannot impose an inference relationship between a verifiable and non-verifiable statement.
+In terms of the natural topology, these correspond to open sets, therefore, by definition, the map is continuous as it maps open sets to open sets.
+The typical requirement that functions are ``well behaved'', then, is actually physically significant.}
 
@@
 The general theory at this point covers only the topological aspects and needs to be extended to the geometrical/measure theoretical ones.
-The idea should be that diverse aspects, such as geometry, measure theory, probability theory, information theory, can be recovered by characterizing the granularity (i.e.
+Our current thinking is that diverse aspects, such as geometry, measure theory, probability theory, information theory, can be recovered by characterizing the granularity (i.e.
 accuracy) of the statements.
-As each statement corresponds to a Borel set of possibilities, assigning a size to these sets effectively means giving a size to the statements themselves.
+Group theoretical aspects may be recovered
+
+\fact{Granularity as foundational concept} {As each statement corresponds to a Borel set of possibilities, assigning a size to these sets effectively means giving a size to the statements themselves.
 For geometry, two sets have the same volume because they contain the same number of possible spatial points.
 For probability, borrowing idea from Cox probable inference, we would assign probability only on conditional statements, therefore the $P(s_1 | s_2)$ would measure what fraction of the possibilities compatible with $s_2$ are also compatible with $s_1$.
-For information theory, given two statements the logarithm of the ratio of the size of the possibilities corresponds to the number of yes/no questions that are needed to go from one statement to the other, the relative information.
+For information theory, given two statements the logarithm of the ratio of the size of the possibilities corresponds to the number of yes/no questions that are needed to go from one statement to the other, the relative information.}
 
-These ideas still need to be developed in a full fledged theory to see whether they can work all the way through.
+\fact{Recover differential geometry or geometry measure theory} {The idea is to start with finite functions of finite k-dimensional surfaces that have the following properties: they are linear if the surface is broken into disjoint parts (the value for the whole is the sum of the values for the parts) and they respect the limit (the value for the limit surface is the limit of the sequence of values for the surfaces).
+Under these, and possibly others, conditions we expect these functions to be expressed through an integral of local functions (i.e.
+k-forms) over infinitesimal surfaces (i.e.
+k-vectors).}
 
@@
 Specifying the state space means characterizing this trade-off for the specific system in the specific circumstances.}
+
+\fact{Determinism and reversibility is isomorphism}{TODO}
 
@@
 
-[To be completed]
+\fact{Link between quantum hydrodynamics, stochastic mechanics, Bohmian mechanics}{We should be able to reconcile the fact that quantum mechanics can be reframed in these different ways.
+The fact that hydrodynamics-like equations emerge should be linked to the idea that a quantum is a finite size object that can be thought as made of infinitesimal parts, as a fluid.
+Each part, though, is subjected by the Brownian motion associated with the unknowable internal dynamics.
+Stochastic mechanics, then, characterizes those fluctuations while Bohmian mechanics characterizes the average paths.}
+
+\fact{Deriving quantum from classical}{If the above characterizations are understood, we should be able, in line of principle, to take a classical system, couple it with a suitable purely stochastic dynamic and recover quantum mechanics.}
 
@@
 
-\subsection{Other theories}
+\subsection{Further work}
+
+The work still needs to be extended to field theories, classical (electromagnetism and general relativity) and quantum (quantum electrodynamics, quantum chromodynamics, ...).
+Some ideas have been explored but only at a preliminary stage.
 
@@
 
-We haven't yet attempted to work with general relativity as of yet.
-Classical mechanics as derived in this project is already relativistic, in the sense that we derive a metric tensor that is locally Minkowskian.
+\fact{General relativity} {Classical mechanics as derived in this project is already relativistic, in the sense that we derive a metric tensor that is locally Minkowskian.
 However, it is not clear whether that metric has a link to the energy/momentum tensor or it is independent.
+One idea to explore is that the presence of a field changes how states are counted over kinetic momentum (i.e.
+$\{mu^\alpha, mu^\beta\} = F^{\alpha\beta}$).
+It remains to be seen whether the curvature of the metric can be understood in terms of this insight.}
 

### 2019-12-28 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
Updated introduction and citations

@@
 
-The Shannon entropy $-\sum p_i \log p_i$ is well established in such disparate fields as computer science, physics and ecology.
-It is most often introduced in physics through statistical mechanics, and associated with vague notions, such as uncertainty, knowledge, lack of knowledge and disorder.
-These prevent a proper understanding of the concept and often lead to confusion.
+The Shannon entropy $-\sum p_i \log p_i$ is well established in such disparate fields as computer science, physics\cite{Jaynes1957-1, Grandy} and ecology, yet it is still surrounded by a great deal of confusion.\cite{Swendsen, KishFerry} For example, in statistical mechanics it is often introduced and associated with vague notions (such as uncertainty, knowledge, lack of knowledge and disorder) which prevent a proper understanding of the concept.\cite{Styer, Ben-Naim} In this work we concentrate on giving a better characterization of the Shannon entropy and its relation to statistical mechanics by addressing what we believe are the two main problems; we do not discuss the relationship to the thermodynamic entropy, which is separate concept.
 
-This paper aims to give a crisp characterization of the Shannon entropy that is intuitive and precise.
-The general idea is that it measures the variability of the elements within a distribution, which is a general concept applicable to many branches of science, and it is independent of the notion of entropy of thermodynamics and statistical mechanics.
-Therefore we will use the term \emph{``Shannon variability,''} leaving ``entropy'' to the physical concept.
-We will show that the expression $- \sum_i p_i \log p_i$ is the only linear indicator of variability.
-It measures the number of questions one must ask to identify an element of the distribution, linking its use to information theory.
-We will study how the expression works over continuous variables, and show how phase space is special as it leaves the Shannon variability invariant under change of coordinates.
-We will turn to statistical mechanics and apply Shannon variability in different ways to recover connections to the Boltzmann, Gibbs and von Neumann entropies.
+The first problem is that the formula is attached to different concepts that are specific to each field.
+To disentangle this issue we need to provide a characterization of the Shannon entropy that satisfies the following requirements:
+\begin{itemize}
+	\item it is defined in a general setting that is crisp and intuitive
+	\item it is precise enough so that the formula itself can be derived
+	\item it recovers the specific meaning when applied to each field.
+\end{itemize}
+In this paper we show that the idea of variability of the elements within a distribution, how diverse the objects are from each other within the collection, can serve as such a general concept.
+We will therefore use the term \emph{``Shannon variability,''} leaving ``entropy'' to the physical concept.
+We will show that the expression $- \sum_i p_i \log p_i$ is the only linear indicator of variability and it measures the number of questions one must ask to identify an element of the distribution, linking its use to information theory.
+The variability of messages within a communication channel will correspond to its information entropy as used in communication theory; the variability of microstates within a macrostate will correspond to the Gibbs entropy used in statistical mechanics; the variability of the state of a single particle within a microstate will correspond to the Boltzmann entropy.
+
+The second problem is that the Shannon formula presents some peculiarity when applied over continuous variables that is often the source of confusion.
+Namely, it is generally linked to a choice of unit.
+Yet, the physical state spaces, in both classical and quantum mechanics, provide additional structure that make the Shannon variability coordinate-independent.
+This is seldom pointed out, and it is critical to form a proper connection to the Boltzmann, Gibbs and von Neumann entropies.
 
@@
 Therefore we assume little knowledge of the subject from the reader and include all details that we feel are important to properly understand the subject without confusion, including two standard derivations for the formula and examples that are common in computer science but rarely used in physics.
-We will briefly touch on a few of the common misconceptions.
+We will briefly touch on a few of the common misconceptions and show how they are resolved.
 

### 2020-01-13 Articles/Overview/overview.tex
Overview: minor tweaks

@@
 
-\extendedfalse
+\extendedtrue
 
@@
 \author{Gabriele Carcassi, Christine A. Aidala}
-\date{November 2019}
+\date{January 2020}
 
@@
 \begin{abstract}
-	This article gives an overview of the project ``Assumptions of Physics'' which aims to rederive the known laws from a few physically meaningful starting points.
+	This article gives an informal overview of the project ``Assumptions of Physics'' which aims to rederive the known laws from a few physically meaningful starting points.
 It presents the motivations behind the project, a summary of the main findings and current status of the research.
@@
 Each subsection will be fairly self-contained to allow the reader to skim through the document and focus on the parts of interest.
-It will contain a summary section at the beginning and a titled paragraph for each of the main idea or finding.
+It will contain a summary section at the beginning and a titled paragraph for each of the main ideas or findings.
 

### 2020-01-17 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Addressing reviewers comments

@@
 
-Proper understanding of a physical theory means answering one question: why does the theory work?
-And the answer cannot be: because it does experimentally.
-The problem is not that the answer begs the question; it is that no theory works experimentally for all possible systems in all possible settings.
-In this case, there are no ``right" theories, only ones that agree with experiments in a set of circumstances.
-The question, then, should be really taken to mean: what are all the necessary conditions under which the theory works?
-What are the assumptions or abstractions implicit in the model and where do they apply?
+Proper understanding of a physical theory means answering, among others, the question: why does the theory work experimentally?
+Why does the prediction match our data?
+And the answer cannot merely be: because it does.
+The problem is not just that the answer begs the question; it is that no known theory works experimentally for all possible systems in all possible settings.
+Newtonian gravitation does not predict gravitational lensing; general relativity is not expected to hold at the smallest scales.
+There are no ``right" theories in an absolute sense, only ones that agree with experiments in a set of circumstances.
+So it is not clear why in some cases disagreement with experiment would discard the theory altogether while in others it would not.
 
-To us, this is not an abstract intellectual question.
-It is a practical matter: I have a system to study and have a choice of different models (e.g.
+We need to rephrase the question as: why does the theory work experimentally in these cases, but not those others?
+In other words, what is the realm of applicability of our theory?
+If the theory disagrees with experimental data within the realm of expected applicability, then the theory will be either discarded or amended in its applicability.
+In practice, physicists and engineers learn by examples which particular theory is suitable for what system, without a clear-cut line of demarcation.\footnote{To illustrate this with an anecdote: it is not uncommon to find individuals, even physics professors, who will insist that Hamiltonian mechanics is perfectly equivalent to Newtonian mechanics.
+When asked to provide the Hamiltonian for a damped harmonic oscillator, they will right away note that they cannot because it is a dissipative system.
+Then they will typically either admit that the two are different, or that they are still the same but only when they are both applicable.} Ideally, it would be preferable to find a set of necessary conditions that the system needs to satisfy for the theory to be applicable.
+This would give a clear line of demarcation within which empirical adequacy is both expected and justified.
+The description of the system in the theory matches experimental verification because, in those circumstances, the system can be considered to satisfy those conditions.
+If we chose to apply the theory where those conditions were not satisfied, it would not be the theory's fault that it gave incorrect predictions.
+It would be our fault for incorrectly applying it.
+
+To us, this is not just an abstract intellectual question.
+It is a practical matter.
+I have a system to study and have a choice of different models (e.g.
 Hamiltonian mechanics, Newtonian mechanics, quantum mechanics, Markov processes, ...).
 How do I know which one to pick?
+What should I tell a student?
 
-If Hamiltonian mechanics is characterized as describing point-like objects, whose state is described by position and momentum, and that move according to a specified set of equations, it is not clear why it would apply to a planet and to a speck of dust in a vacuum, but not to a speck of dust in a fluid or to an electron.
-The size does not seem to matter and it is not clear why Hamilton's equations are important: why should a point-particle be constrained to follow those equations?
+If we characterize classical Hamiltonian mechanics as describing point-like objects, whose state is described by position and momentum, and that move according to a specified set of equations, it is not clear why it would apply to a planet and to a speck of dust in a vacuum, but not to a speck of dust in a fluid or to an electron.
+The size does not seem to matter while the circumstances seem to.
+Saying that Hamiltonian mechanics describes point-particles does not clarify what physical objects can be described by the theory: it does not help us understand what classical systems are.
+Moreover, why should a point-particle be constrained to follow those equations?
+A damped harmonic oscillator does not.
+Why is Hamiltonian mechanics not applicable there?
+Therefore this characterization does not help us identify the realm of applicability.
 
-If we instead characterize Hamiltonian mechanics as the deterministic and reversible motion of objects reducible to infinitesimal parts, we understand when the model choice is appropriate.
+If we instead characterize classical Hamiltonian mechanics as the deterministic and reversible motion of objects reducible to infinitesimal parts, we understand when the model choice is appropriate.
 Can we pretend that the system we are studying, under the conditions we are studying it, is made of infinitesimal parts that can be studied independently?
@@
 For a planet and speck of dust in a vacuum, yes.
+With this characterization, then, a classical system is one that can be considered to be infinitesimally reducible.
+It is furthermore Hamiltonian if the evolution is deterministic and reversible.\footnote{In the same vein, \cite{AoPPhy1} finds that it is Newtonian if it is purely kinematic (i.e.
+spatial motion identifies the state), it is Lagrangian if it is both Hamiltonian and Newtonian and so on.} The structure of phase space and the equations of motion are a direct consequence of those conditions.
+It is clear what the realm of applicability of the theory is and its empirical validity is justified within that realm.
 
-The characterization of classical mechanics as describing point-particles, then, fails spectacularly at answering the most basic question: what is a classical system?
-Because knowing what a classical system is means knowing when classical mechanics is applicable.
-And saying that classical mechanics describes point-particles does not clarify when classical mechanics is applicable.
-Our characterization, instead, is very straightforward: a classical system is one that can be considered to be infinitesimally reducible.
-It is furthermore Hamiltonian if the evolution is deterministic and reversible.\footnote{In the same vein, \cite{AoPPhy1} finds that it is Newtonian if it is purely kinematic (i.e.
-spatial motion identify the state), it is Lagrangian if it is both Hamiltonian and Newtonian and so on.}
-
-Finally, note that the approach presented leads to both the structure of phase space and the equations of motion.
-There is no freedom to choose other structures.
-With point-particles, instead, it is not clear why, for example, position is not sufficient to specify the state, or why trajectories that do not obey Hamilton's equations must be discarded.
-Therefore the point-particle characterization has far less explanatory power.
+Therefore we prefer the characterization presented here as it gives a set of necessary conditions from which we can rederive the theory.
+These act as a clear demarcation for when the theory is expected to match experimental data (i.e.~its realm of applicability), which we believe is a necessary element for a full understanding of any physical theory.
 

### 2020-01-18 Articles/HamiltonianMechanicsIsConservationOfEntropy/refereeResponse2.tex
Adding referee response

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{dutchcal}
\usepackage{braket}
\usepackage{enumitem}

\usepackage{tikz}
\usepackage{forest}
\usetikzlibrary{trees}
\usetikzlibrary{calc}
\usepackage{calculator}
\usepackage{standalone}

\begin{document}

\section*{Referee response}

\emph{We thank the reviewer for the additional comments and we are pleased two of the previous points were addressed satisfactorily.}
\bigskip

\textbf{The additional section on infinitesimal parts vs point particles also does much to clarify the authors' claim that the former should be the objects of classical mechanics. However, I think the argument presented here is still not entirely convincing: the main line of argument seems to claim that a theory should be judged based on how clearly it defines its own extension; in fact, the authors appear to make the claim that this requirement should supplant any considerations of empirical adequacy. Surely, that is a highly controversial claim and would need more argument than is provided. In particular, it is not clear to me why the fact that no theory is perfectly empirically adequate should mean that empirical adequacy cannot be used to judge the merits of a theory. }

\emph{Supplanting any considerations of empirical adequacy would indeed be a controversial claim, and one we would not agree with. We reworked the whole section, hoping we no longer give that impression.  It now reads:}

\subsection*{Point particles vs infinitesimal parts}

At this point one may still argue that point particles and infinitesimal parts are both abstract concepts and both factually incorrect (i.e. planets are neither points nor made of infinitesimal parts), so why should we prefer one concept over the other?

Proper understanding of a physical theory means answering, among others, the question: why does the theory work experimentally? Why does the prediction match our data? And the answer cannot merely be: because it does. The problem is not just that the answer begs the question; it is that no known theory works experimentally for all possible systems in all possible settings. Newtonian gravitation does not predict gravitational lensing; general relativity is not expected to hold at the smallest scales. There are no ``right" theories in an absolute sense, only ones that agree with experiments in a set of circumstances. So it is not clear why in some cases disagreement with experiment would discard the theory altogether while in others it would not.

We need to rephrase the question as: why does the theory work experimentally in these cases, but not those others? In other words, what is the realm of applicability of our theory? If the theory disagrees with experimental data within the realm of expected applicability, then the theory will be either discarded or amended in its applicability. In practice, physicists and engineers learn by examples which particular theory is suitable for what system, without a clear-cut line of demarcation.\footnote{To illustrate this with an anecdote: it is not uncommon to find individuals, even physics professors, who will insist that Hamiltonian mechanics is perfectly equivalent to Newtonian mechanics. When asked to provide the Hamiltonian for a damped harmonic oscillator, they will right away note that they cannot because it is a dissipative system. Then they will typically either admit that the two are different, or that they are still the same but only when they are both applicable.} Ideally, it would be preferable to find a set of necessary conditions that the system needs to satisfy for the theory to be applicable. This would give a clear line of demarcation within which empirical adequacy is both expected and justified. The description of the system in the theory matches experimental verification because, in those circumstances, the system can be considered to satisfy those conditions. If we chose to apply the theory where those conditions were not satisfied, it would not be the theory's fault that it gave incorrect predictions. It would be our fault for incorrectly applying it.

To us, this is not just an abstract intellectual question. It is a practical matter. I have a system to study and have a choice of different models (e.g. Hamiltonian mechanics, Newtonian mechanics, quantum mechanics, Markov processes, ...). How do I know which one to pick?  What should I tell a student?

If we characterize classical Hamiltonian mechanics as describing point-like objects, whose state is described by position and momentum, and that move according to a specified set of equations, it is not clear why it would apply to a planet and to a speck of dust in a vacuum, but not to a speck of dust in a fluid or to an electron. The size does not seem to matter while the circumstances seem to. Saying that Hamiltonian mechanics describes point-particles does not clarify what physical objects can be described by the theory: it does not help us understand what classical systems are. Moreover, why should a point-particle be constrained to follow those equations? A damped harmonic oscillator does not. Why is Hamiltonian mechanics not applicable there? Therefore this characterization does not help us identify the realm of applicability.

If we instead characterize classical Hamiltonian mechanics as the deterministic and reversible motion of objects reducible to infinitesimal parts, we understand when the model choice is appropriate. Can we pretend that the system we are studying, under the conditions we are studying it, is made of infinitesimal parts that can be studied independently? In the case of an electron the answer is no. In the case of a planet or a speck of dust, if we are studying only the overall motion, then yes. Can we pretend that the evolution is deterministic and reversible? In the case of the speck of dust in a fluid, no. For a planet and speck of dust in a vacuum, yes. With this characterization, then, a classical system is one that can be considered to be infinitesimally reducible. It is furthermore Hamiltonian if the evolution is deterministic and reversible.\footnote{In the same vein, \cite{AoPPhy1} finds that it is Newtonian if it is purely kinematic (i.e. spatial motion identifies the state), it is Lagrangian if it is both Hamiltonian and Newtonian and so on.} The structure of phase space and the equations of motion are a direct consequence of those conditions. It is clear what the realm of applicability of the theory is and its empirical validity is justified within that realm.

Therefore we prefer the characterization presented here as it gives a set of necessary conditions from which we can rederive the theory. These act as a clear demarcation for when the theory is expected to match experimental data (i.e.~its realm of applicability), which we believe is a necessary element for a full understanding of any physical theory.

\bibliographystyle{alpha}

\bibliography{bibliography}{}

\end{document}

### 2020-02-04 Proceedings/2019-Quantum2019/CarcassiQuantum2019ProceedingsArchive.tex
Quantum2019 archive version

--- 2019-12-26 Proceedings/2019-Quantum2019/CarcassiQuantum2019Proceedings.tex
@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Trim Size: 9.75in x 6.5in
-%% Text Area: 8in (include Runningheads) x 5in
-%% ws-ijqi.tex   :   18-6-08
-%% Tex file to use with ws-ijqi.cls written in Latex2E.
-%% The content, structure, format and layout of this style file is the
-%% property of World Scientific Publishing Co. Pte. Ltd.
-%% Copyright 1995, 2002 by World Scientific Publishing Co.
-%% All rights are reserved.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-\newcounter{myctr}
-\def\myitem{\refstepcounter{myctr}\bibfont\noindent\ifnum\themyctr>9\else\phantom{0}\fi\hangindent17pt\themyctr.\enskip}
-\def\myhead#1{\vskip10pt\noindent{\bibfont #1:}\vskip4pt}
+\documentclass{article}
+\usepackage{authblk}
+\usepackage{amsmath,amsfonts}
 
-\documentclass{ws-ijqi}
-\usepackage{hyperref}
-\usepackage[super,sort,compress]{cite}
-\usepackage{url}
 \begin{document}
 
-%%%%%%%%%%%%%%%%%%%%% Publisher's Area please ignore %%%%%%%%%%%%%%
-\catchline{}{}{}{}{}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\title{The fundamental connections between classical Hamiltonian mechanics, quantum mechanics and information entropy}
 
-\title{THE FUNDAMENTAL CONNECTIONS BETWEEN CLASSICAL HAMILTONIAN MECHANICS, QUANTUM MECHANICS \\ AND INFORMATION ENTROPY}
+\author[\dagger]{Gabriele Carcassi}
 
-\author{GABRIELE CARCASSI}
+\author[\dagger]{Christine A. Aidala}
 
-\address{Physics Department, University of Michigan, 450 Church Street\\
+\affil[\dagger]{Physics Department, University of Michigan, 450 Church Street\\
 Ann Arbor, MI 48109-1040,
@@
 
-\author{CHRISTINE A. AIDALA}
-
-\address{Physics Department, University of Michigan, 450 Church Street\\
-Ann Arbor, MI 48109-1040,
-United States\\
-caidala@umich.edu}
-
 \maketitle
-
-\begin{history}
-%\received{Day Month Year}
-%\revised{Day Month Year}
-%\accepted{Day Month Year}
-%\comby{(xxxxxxxxxx)}
-\end{history}
 
@@
 
-\keywords{Information entropy; Hamiltonian mechanics; Quantum mechanics; determinism and reversibility.}
+%\keywords{Information entropy; Hamiltonian mechanics; Quantum mechanics; determinism and reversibility.}
 
@@
 By assumptions here we mean simplifying conditions that our system is supposed to satisfy under the processes under study.
-Our work differs from others, see for example Refs.~\citenum{PhysRevA.84.012311,QLogicReview,Hardy:2001jk,ludwig_hein_2013}, in that the starting points are physical assumptions about the system being studied (e.g.
+Our work differs from others, see for example Refs.~\cite{PhysRevA.84.012311,QLogicReview,Hardy:2001jk,ludwig_hein_2013}, in that the starting points are physical assumptions about the system being studied (e.g.
 whether the parts can be studied or whether the evolution is deterministic) and not information theoretic or abstract mathematical considerations.
@@
 This classical uncertainty relationship is reminiscent of the quantum uncertainty relationship in both the form and the fact that the equality holds for the Gaussian packets.
-The analogy can be made stronger with a few extra conditions as described in Ref.~\citenum{Carc1}.
+The analogy can be made stronger with a few extra conditions as described in Ref.~\cite{Carc1}.
 What we want to show here is the connection between this uncertainty relationship, information entropy, deterministic and reversible evolution and the difference in reducibility between a classical and a quantum system.
@@
 
- Funding for this work was provided in part by the MCubed program of theUniversity of Michigan.
+Funding for this work was provided in part by the MCubed program of the University of Michigan.
 
-\renewcommand\bibname{References}
+%\renewcommand\bibname{References}
 
-\bibliographystyle{ws-ijqi}
+\bibliographystyle{plain}
 \bibliography{bibliography}

### 2020-02-04 Proceedings/2020-Foundations2020/Abstract.tex
Abstract for Foundations2020

\documentclass{article}
\usepackage{authblk}
\usepackage{amsmath,amsfonts}
\usepackage[margin=1.5in]{geometry}

\usepackage{cite}

\begin{document}

\title{Insights from Assumptions of Physics}

\author[1]{Gabriele Carcassi}

\author[1]{Christine A. Aidala}

\author[2]{David J. Baker}

\author[1]{Kai Sun}

\affil[1]{Physics Department, University of Michigan}
\affil[2]{Philosophy Department, University of Michigan}

\maketitle

\section*{Extended Abstract}
Assumptions of Physics\cite{A,B,C,D} is a project that aims to find a minimal set of physical assumptions from which the basic laws can be derived. We do so by combining ideas not only from different branches of physics, but also advances in other disciplines (e.g. mathematics, theoretical computer science) that have so far not reached mainstream physics. The general idea is to construct a single mathematical framework with structures that apply in the general case, which get further specialized when taking further assumptions. Experimental verifiability, a fundamental requirement in science, already forces a particular mathematical structure; the notion of states and processes imposes more constraints; the additional notion of infinitesimal reducibility (i.e.~state of the whole system is equivalent to its infinitesimal parts) imposes enough structure to recover classical phase space.

In this talk we will present a few key insights with which our work on reformulating different scientific theories forced us to come to terms. We believe these are of general interest to those who work on the foundations of current theories or are working on the development of new, more general ones. A couple of examples follow.

It is common nowadays  to develop or present a theory by starting with its theoretical content (e.g.~relationship between perfectly defined elements and their infinitesimally precise description) and then recovering the empirical content (e.g.~the finite precision measurements and relationships between them). What we have found is that it is more fruitful to proceed the other way. This can be done because of two results. The first is that the theoretical content can be recovered from the empirical one - mathematically, the Heyting algebra that describes the experimentally verifiable statements can be used to recover the points and their topology. The second is that an inference relationship among verifiable statements is equivalent to a causal relationship. Therefore the theoretical content is effectively a more compact way to describe the empirical content.

A more predominant role of the experimental verification leads us to a different perspective on reductionism, the description of a system through its constituents. This is typically done by describing the smallest systems and then aggregating them. We find it more fruitful to start from the overall system and ask what are the requirements to be able to identify and describe the parts, which is more in line with what happens experimentally. For example, when assuming that a system is infinitesimally reducible (see above) the structure of classical phase space is recovered; on the other hand, if the system is assumed irreducible (i.e.~the state of the overall system does not tell us anything about the parts, its internal dynamics) the state space of quantum mechanics is recovered. These assumptions are contingent upon both the system and the processes under study. For example, a proton can be adequately described as a single quantum system when the process (e.g.~diffraction and interference experiments) does not probe the internal dynamics. If it does (e.g. deep inelastic scattering) the assumption of irreducibility fails. This suggests to take a top-down, more practical epistemic attitude toward reductionism (i.e.~is the given process sensitive to the internal dynamics of the system) instead of a bottom-up, more ontological attitude (i.e.~the system is made of these parts). A final theory will then need to explain when and how the internal dynamics of each system becomes accessible.

Not only does the state of a system depend on the accessibility of its internal dynamics, but it additionally depends on its interaction with the environment at its boundary. The position of a plastic ball makes sense on earth at 20 Celsius. On the surface of the sun at 5500 Celsius there would be no ball to speak of. When studying a system, the system-environment interaction has to be such that the system remains itself - mathematically the state space of the system is a symmetry of the system-environment interaction. While these considerations are included in the theorical treatment of condensed matter, statistical mechanics and thermodynamics, in quantum mechanics and other foundational theories they are often thought as mere experimental details. For example, since experimentally producing and maintaining entangled or coherent states requires very careful control of the system-environment interaction, it is misleading to attribute the behavior solely to the system. A final theory will need to take into account the system-environment boundary and explain how it inhibits or enhances the possible dynamics of the system.

\bibliographystyle{simple}
\begin{thebibliography}{4}

\bibitem{A} G. Carcassi, C. A. Aidala, D. J. Baker, L. Bieri: From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics. Journal of Physics Communications 2-4 (2018)

\bibitem{B} C. A. Aidala, G. Carcassi, M. J. Greenfield: Topology and experimental distinguishability. Topology Proceedings 54 (2019)

\bibitem{C} G. Carcassi, C. A. Aidala: The fundamental connections between classical Hamiltonian mechanics, quantum mechanics and information entropy. Accepted in International Journal of Quantum Information (2020)

\bibitem{D} http://assumptionsofphysics.org

\end{thebibliography}

\end{document}

### 2020-02-04 Proceedings/2019-ICNFP/Carcassi2019ICNFPProceedings.tex
First pass at ICNFP2019 proceedings

\documentclass[12pt]{iopart}
\newcommand{\gguide}{{\it Preparing graphics for IOP Publishing journals}}
%Uncomment next line if AMS fonts required
\usepackage{iopams}  
\begin{document}

\title{Space-time structure may be topological and not geometrical}

\author{Gabriele Carcassi, Christine A. Aidala}

\address{Physics Department, University of Michigan, 450 Church Street\\
	Ann Arbor, MI 48109-1040,
	United States}
\ead{carcassi@umich.edu}
\vspace{10pt}
\begin{indented}
\item[]February 2019
\end{indented}

\begin{abstract}
In a previous effort we have created a framework that explains why topological structures naturally arise within a scientific theory; namely, they capture the requirements of experimental verification. This is particularly interesting because topological structures are at the foundation of geometrical structures, which play a fundamental role within modern mathematical physics. In this talk we will show a set of necessary and sufficient conditions under which those topological structures lead to real quantities and manifolds, which are a typical requirement for geometry. These conditions will provide a physically meaningful procedure that is the physical counter-part of the use of Dedekind cuts in mathematics. We then show that those conditions are unlikely to be met at Planck scale, leading to a breakdown of the concept of ordering. This would indicate that the mathematical structures required to describe space-time at that scale, while still topological, may not be geometrical.
\end{abstract}

%
% Uncomment for keywords
%\vspace{2pc}
\noindent{\it Keywords}: Foundations of physics, space-time structure, topology

%
% Uncomment for Submitted to journal title message
\submitto{\PS}
%
% Uncomment if a separate title page is required
%\maketitle
% 
% For two-column output uncomment the next line and choose [10pt] rather than [12pt] in the \documentclass declaration
%\ioptwocol
%

\section{Introduction}

\section{Elements of topology and its link to experimental verifiability}

The first thing we need to establish is the link between topology and experimental verifiability. Topology is, knowingly or unknowingly, widely used in physics as it is the foundation of many other tools, such as differential geometry and Lie groups. Most mathematical structures used in physics are topological spaces. Therefore it is crucial to understand what physical content do they capture.\footnote{In computer science, the link between topology and computability is already accepted though not widely known. There is a link between those concepts and the ones presented here, but, for brevity, we are not going to expand on it. The gist is that any computation device is also a physical system, and the output of a computation can be experimentally verified (i.e. we can read it).}

The formal definition of a topology is the following: given a set $X$, a topology $\mathsf{T}$ is a collection of subsets of $X$ that is closed under finite intersection and arbitrary (infinite) union.\footnote{Technically, it also must contain the full set $X$ and the empty set $\emptyset$. This does not play an important role.} This is similar to other algebraic structures, like groups, where you have elements and operations that return other elements of the same structure. Here the elements are subsets of $X$ and the operations are set operations. The formal definition is very abstract and does not have an apparent connection to physics.

Seemingly unrelated, consider a ``verifiable statement'': an assertion that can be verified to be true experimentally. For example, ``the mass of the neutrino is $0.05 \pm 0.005$ eV" or ``the electron is a negatively charged particle''. In general these statements do not perfectly identify a single possible case, rather they identify a set of possible cases. If $X$ is the set of all possible cases, each verifiable statement is associated with a set $U \subseteq X$ and the statement can be re-expressed as ``$x$ is in $U$''. In our first example, $x$ is the mass of the neutrino and $U$ is the set of possible values from $0.045$ and $0.055$.

Now we note that not all subsets of $X$ correspond to a verifiable statement. For example, ``the mass of the neutrino is exactly $0.05$ eV'' is not verifiable because we cannot perform a measurement with infinite precision. We call $U \subseteq X$ a verifiable set if the statement ``$x$ is in $U$'' is, at least in line of principle, something we can verify experimentally. Let $\mathsf{T}$ be the collection of all verifiable sets. The idea is that $\mathsf{T}$ is a topology.

To convince ourselves of that, consider the following. If we have a finite collection of verifiable sets $\{U_i\}_{i=1}^n$, then the intersection corresponds to the logical AND (i.e. the conjunction) of all the statements ``$x$ is in $U_i$''. We can test the intersection simply by testing each assertion one at a time, and therefore $\{U_i\}_{i=1}^n$ is a verifiable statement. If the collection were infinite, though, we would have to go through infinitely many tests to verify the logical AND, so we would never terminate. Therefore, in general, the infinite intersection of verifiable sets is not verifiable.

On the other hand, the union of verifiable sets would correspond to the logical OR (i.e. disjunction) of all the associated statements. In this case, as long as one statement is verified, the OR is verified and we can terminate. Therefore we do not care how many statements are there after the one that was verified, so the infinite union of verifiable sets is verifiable. Technically, topologies are closed under arbitrary unions, not just countable. Yet, because we cannot test more than countably many statements, the set of all verifiable statements must be constructable for a countable set of statements, or we would not be able to fully explore the space even with unlimited time. Therefore the topology $\mathsf{T}$ must have a countable basis, which means all arbitrary unions are countable unions.

We can make this more concrete by looking at the standard topology on the real numbers $\mathbb{R}$, which contains all the open intervals $(a,b)$ where $a,b \in \mathbb{R}\cup\{-\infty, +\infty\}$. These, in fact, correspond to all the finite precision measurements. It will also contains their unions. For example ``the absolute value of the charge of the electron is $1.6 \pm 0.005\;10^{-19}$ C'' would correspond to ``$x$ is in $(-1.65 \;10^{-19}, 1.55 \;10^{-19})\cup(1.55 \;10^{-19},1.65 \;10^{-19})$. Singletons, sets with a single value, are not part of the topology and in fact we cannot verify them experimentally. Note that the topology can be generated by the set of all rational intervals, which is countable, since the infinite unions will allow us to construct limits, which correspond to the intervals of the reals.

A topology, then, is not just some mathematically abstract construction that happens to be useful in physics. It captures the way that we can experimentally distinguish a set of possible cases. The main point is that statements have a binary logic in terms of TRUE/FALSE, while experimental tests have a ternary logic in terms of SUCCESS/FAILURE/UNDEFINED. The topology is generally used to keep track of those differences. It makes sense, then, that topologies are so pervasive among the mathematical structures used in physics: experimental verifiability is at the heart of science.

Note that we will call verifiable sets, instead of open sets, the sets within the topology. We will also call falsifiable sets, instead of closed sets, the complement of verifiable sets. It will make our work more connected to this physical notion and therefore a little bit more intuitive, especially to those who are not deeply familiar with point-set topology.

The question we want to answer is the following: how are the real numbers constructed in physics? What are the requirements on our experimental apparatus such that those quantities can be measured? Can we expect those requirements to hold when we go to ever smaller scale? If not, how do they break down?

Since we established a link between topology and experimental verifiability, we can pose that questions in a way that is both mathematically and physically precise. Suppose we have a set of possible cases $X$, which we will call possibilities, and a way to distinguish them experimentally, which means we have a topology $\mathsf{T}$ that captures all our verifiable sets. What are the requirements on $(X, \mathsf{T})$ such that it is homeomorphic (i.e. topologically equivalent) to the set of the real numbers with standard topology?

\section{References as the starting point for measurement scales}

The first thing we need to develop is an abstract model to represent how quantities are measured experimentally. Let us first gather some requirements from the physics.

We start from the notion of a reference, such as a mark on a ruler or the tick of a clock. A reference, then is a physical object that allows us to distinguish between three cases: a before, an on and an after the reference. That is, a point can be before, on or after the mark on a ruler; an event can happen before, on or after the tick of a clock. We take references to be the basic conceptual element upon which quantitative measuring devices are constructed in practice.

In general, the three cases a reference defines are not mutually exclusive: an object can extend before, on and after the mark; an event can start before and end after the tick of a clock. In fact, if the object is both before and after, we expect it to be on as well. In all cases, the object should be found at least in one of the three cases.

If a reference is at the end of the range, then it will not have an after or a before. Yet, a reference is a physical object and needs to take some space, therefore there will always be an on case.

We should also note that the before and after cases are easier to test experimentally. When comparing two weights on a scale, for example, it is easy to tell when one is greater than the other. If they are very close, we can only typically say that they are closer than a certain threshold. This is also something we will have to take into account.

If $(X, \mathsf{T})$ is the set of physically distinguishable cases, we can formalize a reference as a triplet $r = (B, O, A)$ of three subsets $B, O, A \subseteq X$. Respectively, they will represent the cases in which the object is before, on or after the references. We can capture the rest of the requirements with the following:
\begin{itemize}
	\item $B \cap A \subseteq O$ (all possibilities in which the object is found before and after, the object is also found on)
	\item $B \cup O \cup A = X$ (before/on/after cover all possibilities)
	\item $O \neq \emptyset$ (the reference itself must be found somewhere)
	\item $B, A \in \mathsf{T}$ (before and after are verifiable)
\end{itemize}
Note that we require only before and after to be experimentally verifiable, leaving the on case unspecified. It turns out this is sufficient and avoids the problem of testing for perfect equality.

A measuring device will be composed of several references. For a continuous quantities, we can intuitively understand that we are assuming that, though we always have a finite set with finite precision, we assume we can refine our instruments with ever greater precision. Our work is to make this intuitive notion precise.

\section{Elements of order theory}

There are many ways to mathematically characterize the real numbers. The one we are interested is in terms of their order, instead of operations like addition or multiplication, as it is the one that imposes the least requirements and it is enough to identify the topology.\footnote{Note how all transformations that preserve addition or multiplication must preserve the ordering, while the converse is not true. For example, non-linear monotonic transformations preserve the ordering and the topology while they do not preserve the group structure.} Let us review, then, a few key concepts of order theory, which is the branch of mathematics that studies ordered sets and their properties.

A partial order $\leq : X \times X \to \mathbb{B}$ is a binary relationship that is reflexive (i.e. $a \leq a$), antisymmetric (i.e. if $a\leq b$ and $b \leq a$ then $a = a$) and transitive (i.e. if $a \leq b$ and $b \leq c$ then $a \leq c$). A total order, or linear order, is a partial order such that any two elements are comparable (i.e. at least $a \leq b$ or $b \leq a$).

Given an ordered set $(X, \leq)$ we can construct the order topology in the following way. Take all sets of the form $(a, \infty) = \{x \in X \, | \, a < x\} \;,\; (-\infty, b) = \{x \in X \, | \, x < b\}$. This is a basis for the topology. Then take all the sets that can be generated from the basis through finite intersection and arbitrary union. This will be the order topology. Both the integers and the reals are totally ordered sets with their standard $\leq$ relationship. Their respective order topology correspond to their standard topology. Note how each set within the basis corresponds a statement like ``x is after a'' or ``x is before b''. Note how these, already, are very similar to the before and after cases we introduced in the previous section.

Like groups or topological spaces, two ordered sets are isomorphic if their have equivalent structure. That is, two ordered sets are isomorphic if there is a bijection that preserves the ordering. If two ordered set are isomorphic, then they will have the same order topology and vice-versa. So the ordering identifies the topology of an ordered set and vice-versa.

In light of this, we should break our problem into two parts. First understand what are the requirements on references such that the set of possibilities $X$ is totally ordered and its topology is the order topology. Then what are the additional requirements to be ordered like the real numbers or the integers. It turns out that most of the work lies in the first part, in recovering the linear order.

\section{Experimental requirements for construct real valued quantities}

\subsection{Strict references}

As we saw, in general a reference defines three cases (before/on/after) that may not be mutually exclusive. We say a reference is strict if they are. That is, a reference $r= (B, O, A)$ is strict if $B \cap O = \emptyset$, $O \cap A = \emptyset$ and $B \cap A = \emptyset$. Technically, we would just need that before and after are disjoint ($B \cap A = \emptyset$) and then redefine on as what remains ($O = X / (A \cup B)$). That is, as long as before and after are separate cases, we can redefine on as ``not before and not after''.

The order topology allows us to construct strict references. If we take $a,b \in \mathbb{R}$ such that $b<a$, we have $( (-\infty, b), [b, a], (a, +\infty) )$ that represents a reference that physically extends from $b$ to $a$. Mathematically, we can see that the extent of the reference is represented by a falsifiable (i.e. closed). Conversely, if we want to be able to construct the order topology from references, then, those will need to be strict. Intuitively if we can experimentally define an order on all possibilities, then, given two distinct ones, we must be able to find references that are able to confirm unambiguously that one is before the other. Therefore we must be able to tell before and after apart.

What does requirement entails in practice? References for quantities that we count (i.e. discrete quantities) can always be made strict: either you have $n$ elements, less than $n$ or more than $n$. The reason is that the quantity does not really have an extent and the possible values are well separated. Quantities over which we have an extent, like space and time, are another matter. If the extent of the object being measured is much smaller than the extent of the reference, then the object cannot be found both before and after. In those regimes we can treat the references as strict. But this is an idealization that will break down if the extent of the references is comparable to the extent of what is being measured.

\subsection{Aligned references}

As we put references together, we must be sure they are related to each other in the correct way if we want to end up with a linear ordering. Intuitively, if we mix reference of horizontal position with vertical positions we will not end up with a linear ordering.

To define what it means for two references to be pointing in the same direction, take two values $b_1, b_2 \in \mathbb{R}$. Suppose that $b_1 \leq b_2$. Then if we find that our object is before $b_1$ then it will also be before $b_2$. In terms of sets, $(-\infty, b_1) \subseteq (-\infty, b_2)$. In a linear order, the idea that one point is before the other can be translate in terms of set inclusion. Alignment between references, then, can be defined by requiring that the before and after statement have an inclusion relationship.

We also note that the negation of ``$x$ is greater than $a$ is ``$x$ is less or equal to $a$''. In terms of sets, $(a, +\infty)^C = (-\infty, a]$. For a strict reference extending from $a$ to $b$, we will have $(-\infty, b) \subseteq (-\infty, a]$.

We will say that two references are aligned if their edges can be ordered. Mathematically two references $r_1 = (B_1, O_1, A_1)$ and $r_2 = (B_2, O_2, A_2)$ are aligned if the sets $B_1, B_2, A_1^C, A_2^C$ can be ordered by inclusion. For example, if $B_1 \subset A_1^C \subset B_2 \subset A_2^C$ then we will be in the case where $r_1$ is completely before $r_2$. If $B_1 \subset B_2 \subset A_1^C \subset A_2^C$ then the two references are overlapping. Alignment of references is another necessary condition if we want to reconstruct and ordering that is linear.

What does this requirement entails in practice? It means that there are clear fixed ordering relationships between the references. For example, every time something is before one reference it will also be after the other. It means we are perfectly able to prepare, control and identify our references.

\subsection{Refinable references}

Strict and aligned references allow us to define ordering between sets of possibilities. We need an additional condition to make sure we have enough references at an appropriate resolution to be able to order the possibilities themselves. That is, to order all singletons, the sets with only one possibility.

The general idea is we must be able to fill in the gaps and break apart overlapping regions. For example, if we have two references that can have something in between, then we must be able to put a reference between them. That is, if $r_1 = (B_1, O_1, A_1)$ and $r_2 = (B_2, O_2, A_2)$ are such that $A_1 \cap B_2 \neq \emptyset$, then we can find a reference $r_3 = (B_3, O_3, A_3)$ such that $O_3 \subseteq A_1 \cap B_2$. On the other hand, if the extent of one reference is within the extent of the other, then we need to be able to find a reference that covers another part. That is, if $O_2 \subset O_1$ then we can find another reference such that $O_3 \subset O_1$ and $O_2 \cap O_3 = \emptyset$. If a set of references have this property, then we say it is refinable: the references can be refined to non overlapping references that cover all the possibilities.

What does this requirement entails in practice? It means that we can place a reference wherever we want and can shrink their extent to the point they occupy only one possibility.

\subsection{Linear order}

It can be demonstrated that a set of experimentally distinguishable possibilities is linearly ordered if and only if its topology can be generated by a set of refinable, aligned strict references. That is, if we have an order topology, we can construct a set of references that have those properties, and if we have a set of references that have those properties, the order topology can be generated by finite intersection and arbitrary union from the before and after sets of the references.

The idea is that, under those conditions, we can find references that are so fine that they extend over only one possibility. Therefore we have a one to one correspondence between experimentally distinguishable cases and the finest references. Since all references are aligned and strict, two different references must be sequential, one before the other. This order corresponds to the one that defines the order topology.

\subsection{Discrete (integer) order}

The order of the integers can be characterized as the one that does not have a minimum, does not have a maximum and, taken any two elements, there are finitely many ones.

If we have an ordered set with the same characteristics, we can put it in a bijective correspondence with the integers that preserves the order. Roughly, one can proceed this way. Take one element and arbitrarily label it zero. Since between two elements there are only finitely many, each element will have a successor and a predecessor. Label one the successor of zero and label minus one the predecessor of zero. Then label two the successor of one, and so on. Since between zero and any other element there are only finitely many other elements, at some point we will reach all the elements.

If we have a set of refinable aligned strict references, then the only additional requirement needed to recover the integers is that between two references there are only finitely many references. The finest possible references, which will correspond to the possibilities of the space, can then be labeled by the integers. The possibilities are ordered like the integers.

\subsection{Continuous (real) order}

The order of the reals can be characterized as one that is dense (i.e. between two elements there is always another one), complete (i.e. contains all the limit points), has a countable dense subset (i.e. between two real numbers we can always find a rational, which is a countable set), has no minimum or maximum.

If we have an ordered set $X$ with the same characteristics, we can put it in a bijective correspondence with the reals that preserves the order. Roughly, one can proceed this way. We start by noticing that we have a countable dense subset $Q$ and, as Cantor showed, it can be put into correspondence with the rationals. The set $Q$ is dense in the original set $X$ and the original set $X$ is complete. Therefore the $X$ is the completion of $Q$. Another theorem in order theorem states that the completion of any order is unique. Since the real numbers are the completion of the rationals, and $Q$ has the same order of the rational, then $X$ has the same order of the reals.

If we have a set of refinable aligned strict references, then the only additional requirement needed to recover the reals is that between two references there is always another reference. That is, we are only imposing that the references are dense. The completion and the countable dense subset are a consequence of the more general topological requirements. The inclusion of all the limit points will come from the infinite union allowed by the topology while the dense subset is associated with the countable base, which is a requirement for out topology to be physically meaningful.

\section{Breakdown of ordering and of geometry}

In the previous sections we have identified the experimental requirements needed to be able to operationally define continuous quantities. These are:
\begin{enumerate}
	\item the ability to compare the quantity (i.e. position, time, mass, ...) with references such that we can say whether the object is before or after each one
	\item the three different cases, before/on/after, must be mutually exclusive
	\item the references must be aligned, we must be able to place them in a line
	\item we must be able to find enough references such that overlaps can be avoided and all the possible values can be covered
	\item between two references we can always place another
\end{enumerate}
Under these conditions, we can assign a real number to the finest possible references, which will correspond to all possible distinguishable cases. Higher dimension can be reconstructed by combining sets of references for different quantities, which leads to the familiar structure of a manifold.

We want to stress that these requirements are not properties of the quantity being measured. They are operational requirements. They correspond to the mundane act of constructing a ruler, a calorimeter, a timing system, or any other experimental device.

We are not, therefore, positing that there already exists a continuous quantity to be found. We are asking, under what conditions we can label the outcomes of our measurement scheme with a continuous quantity. This in line with the perspective of considering a physical theory, including quantum theory, as dealing with the output of measurements. For example, Wheeler's commented: ``It is wrong, moreover, to regard this or that physical quantity as sitting out therewith this or that numerical value''. What we are doing is go at a more fundamental level to understand this works in the continuous case, and how can it be modeled in a rigorous say.

Now that we understand what the requirements are we can ask: are these requirement tenable? Or are they necessarily idealizations?

Note that, with the exception of (v), all other requirements are needed for ordering itself. That is, the failure of any one of the first four means that ordered quantities are not physically meaningful. Which means we cannot substitute the real number with the rationals or the integers, as these are themselves ordered. Note that all geometrical structure are based on some notion of numeric distance, therefore if order fails geometrical structure fail as well. We are not going to be able to define metric tensors, symplectic forms or even a differentiable structure.

As such, the bar to show that, at a fine level, space-time does not posses a geometrical structure is very low. One simply has to instill the doubt that any of (i-iv) is untenable. Here we give a few angles one may take. We will concentrate on measurements of spatial position. Similar arguments can be made for time as well.

For (ii), we saw that references are objects that have, in general, an extent in space. A reference can be considered strict if the extent if the extent of the object measured is much smaller than the extent of the reference. For example, the top level of a liquid in a jar is thinner than the marks used to measure the content. Ultimately, however, the finest constituents will also be the finest objects that we can use as finest references. Therefore, if we assume that we can make finer and finer references, at some point we will find the extent of the references comparable to the extent of the objects we are measuring. This is also true for fundamental particles, where the extent of the system correspond to the extent of the wave function: if the wavefunctions overlap, one particle is not clearly before/on/after the other. Similar arguments apply for excitations of quantum fields.

For (iii), the requirement for two references to be aligned means, for example, that we must always be sure which one is before the other. This means we must have a way to tell the references apart. This is problematic if the fundamental objects are ultimately indistinguishable. We cannot order what we cannot distinguish.

For (iv), the requirement for refinement means we can always prepare finer and finer references. This is generally thought to not be possible once length scales are comparable to the Planck length.

We are not interested here to expand on these arguments or pick ones that are universally accepted. That will be the scope for future work. Here we simply want to show that is relatively easy to formulate reasonable arguments. As long as one finds reasonable that one of (i-iv) fails, then our point follows: space-time structure at a fine scale will not be geometrical, will not be ordered, but it will yet be topological, we will have some way to measure what can physically be well defined.

\section{Conclusion}

It is not clear what topological space should be used instead of a manifold. There are

The simplest solution would be to find a topological space

We leave

Naturally, this opens the question of what topological spaces would be appropriate in those regimes. One can imagine to try and construct a topological space that is not ordered at a fine scale but is ordered at a large scale. A difficulty here is that typically the notion of distance is absent from topological spaces.

There is, however, a much more problematic prospect. General relativity tells us that the geometry depends on the content of the space. What is in the space affects the notion of distance. It would not be far fetch to assume that what is in the space affects what can be distinguished experimentally. This would mean that the topology itself would also depend on the content of the space. In that case, we would not have a fixed topology, in the same way that we do not have a fixed geometry in general relativity. Conceptually, we would need to define some sort of topological equation. Though the premise seems reasonable enough, it is not clear to us whether appropriate mathematical tools for such a line of inquiry currently exist.

\section{Conclusion}

\section*{Acknowledgments}

\end{document}

### 2020-02-14 Proceedings/2019-ICNFP/Carcassi2019ICNFPProceedings.tex
First pass at ICNFP proceeding

@@
 
+Find a few papers to cite that wonder about the structure of space time.
+
+In our project, Assumptions of Physics, the idea is to find a minimal set of assumptions from which the different basic theories can be derived.
+The idea is that, by doing this as formally as possible, we are forced to specify all our implicit starting points and inform us to what can be done different.
+While other approaches start with the same premise, we typically start at a much lower level.
+For example, we do not assume probability spaces which are already constructions upon sigma-algebras and measures, and sigma-algebras are typically constructed from the topology of the space.
+We aim to recover those as well.
+This is particular important if one is interested in the structure of space-time, since its topology and sigma-algebra are the most foundational aspects.
+
+The other difference is that our approach wants to start with primitives that are necessary to do physics.
+Our basic building block is the notion of a verifiable statement: an assertion for which an experimental test is available that would confirm, in finite time, that the assertion is true.
+As one can imagine, without such a notion, we would have no experimental verification, which means no possibility of doing science.
+
+In previous work, we have shown the link with this idea to the mathematical framework of topology.
+That is, in physics, the topology captures what is experimentally verifiable (i.e.
+$U$ is an open set if and only if ``$x$ is $U$'' is experimentally verifiable).
+For real valued quantities, the verifiable statement correspond to open intervals to signify that our measurement have always finite precision.
+The fact that a continuous quantity has precisely a given real value (e.g.
+the length of the side is exactly 1 meter and the diagonal is exactly $\sqrt{2}$ meters) is not experimentally verifiable.
+This gives new insight to the notion of topological continuity and it tells why, in very general terms, functions \textbf{must} be well behaved in physics.
+It is not a matter of convenience: if they are not they would break experimental verifiability.
+
+This idea also clarifies the division between topology and geometry.
+Topological construct are more primitive than geometrical ones because they have no notion of size.
+Physically, it means that before we can assign distances between two points, we must be able to distinguish them, to tell them apart.
+The topology, then, comes first, as it tells us if and how the points are distinguished experimentally.
+This insight also tells us why different spaces, though topologically equivalent, have different geometry.
+In everyday space, the distance along any direction can be measured in the same unit, say meters.
+The geometry is Riemmanian.
+On phase space, within a degree of freedom, units of position cannot be compared to unit of momentums, therefore we cannot define angles.
+Yet, areas of a degree of freedom are proportional to the number of possible configurations.
+The fact that we can quantify areas for each degree of freedom is what gives us symplectic geometry.
+If we imagine the space of all possible blood work results, however, we have neither notions, and therefore we simply have a manifold that is not geometric.
+
+For those interested in the ultimate structure of space-time these insights naturally lead to the following line of inquiry.
+Before we want to understand whether space-time is geometrically truly a Riemmanian manifold, we need to understand whether space-time is topologically truly a manifold.
+A manifold, simply put, is a set of possible cases $X$ that can be experimentally identified by a set of real valued.
+So, the question is: when can a set of possible cases, or possibilities for short, be experimentally identified by a real value?
+To be clear, the question is not when the results of our measurements are real numbers.
+The answer to this is simple: never.
+Our measurements are always finite precision.
+The question is when the cases we want to be able to distinguish are parameterized by the real line.
+
+Even though this work is mathematically very technical and abstract, we strive to have a well understood map to more tangible and physical concepts: that is one of the main goals in our work.
+Therefore let us first give a summary of the results in physics terms, leaving the details to subsequent sections.
+
+\subsection{Summary of results}
+
+% We can cite others from wikipedi
+%F. Faltin, N. Metropolis, B. Ross and G.-C. Rota. The real numbers as a wreath product Advances in Mathematics, 16(1975), 278–304.
+Mathematics offers many constructions of the real numbers from more primitive notions.
+Common techniques include completion through Cauchy sequences, using Dedekind cuts or adding constraints over an algebraic fields.
+In fact, Faltin et al.
+(cite) says:
+\begin{quote}
+Few mathematical structures have undergone as many revisions or have been presented in as many guises as the real numbers.
+Every generation re-examines the reals in the light of its values and mathematical objectives.
+\end{quote}
+Our present purpose is one such re-examination in the light of how these quantities are defined through experiments.
+
+If we think at how quantities are measured, the general idea is that we are able to define references with preset amounts, and then compare our object to several references to find the ones that bound the value from above and below.
+For example, a ruler is a series of marks and measuring the position with it means finding the closest one.
+A clock is a series of ticks and measuring time means noting the tick right before the event.
+A balance scale compares the weight of an object to a few known ones.
+Intuitively, we can understand that a quantity is continuous if, in line of principle, we can prepare references ever closer to each other.
+Our aim is to make this notion precise, so we have a practical/operational definition of what we mean experimentally by a continuous quantity.
+
+A single reference is something that allows us to distinguish a before and an after, a smaller and a greater, a higher and a lower.
+We can formalize a reference with our more basic notion of a verifiable statement: a reference gives us two verifiable statements ``the objects is before this reference'' and ``the object is after this reference''.
+This allows to recast ordering relationship in terms of logical relationship for example, if reference A is after reference B, then if ``the object is before reference B'' is true, ``the object is before reference A'' must also be true.
+And these logical relationships remain conceptually the same regardless of whether ``before'' means ``before in time'' or ``before along a spatial direction''.
+These are the sort of basic ideas and language we use within our framework, and from these we work out a set of necessary and sufficient conditions that a set of references needs to satisfy so that they define a set of possible distinguishable cases, or possibilities for short, that is ordered like the real line.
+
+The biggest obstacle in this undertaking, and possibly in fully appreciating it, is that our standard intuition traps us in circular arguments.
+For example, one issue is that references have an extent: they occupy some space and so do objects.
+If we are to measure the position of an object, then, it may extend before, on and after the mark of a ruler.
+In these cases we typically note the position of the beginning of the object and an end of the object.
+This implicitly relies on the fact that the resolution of our eyes is higher than that of the ruler, and that we can independently recognize parts of the object, namely the beginning and end.
+Identifying parts over a higher resolution correspond, in the end, to new references.
+So, assuming this can always happen means implicitly assuming what we are tasked to derive.
+
+Unfortunately, thinking visually and drawing pictures is inherently misleading because everything we draw is inherently ordered.
+It confuses more than it helps, so much so that we progressed only when we stopped trying to reason with pictures and concentrated on just the boolean logic of our statements in truth tables.
+In that setting, the only elements we are allowed to define are those simple logical relationships we described before: if the object is before A, then it is not after B and it is before C.
+All properties and relationships between references, then, must be defined in that manner.
+
+Using that strategy, we reached two major findings.
+The first major finding is that \emph{most of the conditions are not required by the real numbers specifically, but they are required by ordered quantities in general}.
+That is, the set of reference have to satisfy a few conditions just to be sure that the cases they distinguish experimentally are ordered (i.e.
+can be thought to be all being one after the other).
+The conditions are the following.
+\begin{itemize}
+	\item The references must be strict, meaning an object is either exclusively before, on or after the reference.
+No overlaps are allowed.
+In practice, it means that the extent of the object being measured is always much smaller than the extent of the reference, therefore it will always be considered wholly before, on or after the reference.
+	\item The references must be aligned, meaning their before and after statements identify incremental regions.
+In practice, this means we can always distinguish between them and they remain with constant before/after relationships (e.g.
+they do not fluctuate).
+	\item The references must be refinable, meaning we can always resolve overlaps and fill in the whole space.
+For example, if an object can be between two references, then we must be able to put a reference as well.
+If we have two references that overlap, then we must be able to find other references that fit within that overlap, to tell what is before and after.
+\end{itemize}
+These are the conditions one must require to have any linear order, either discrete (which can be labeled by integers), continuous (which can be labeled by real numbers) or any other.
+The divergence between real and integer quantities lies in how many references one can put in between two references.
+In the integer case, we can only fit finitely many references.
+In the real case, we can fit infinitely many.
+
+The second finding is that \emph{the inability to distinguish below a certain scale most likely invalidates those conditions, leading to a breakdown in ordering.} In other words, it is not just that the points of our space can't be labeled by real numbers: they cannot be ordered and therefore cannot be labeled by any ordered quantity.
+To obtain that result, one must simply argue that at least one of the three conditions fails.
+For example, one could argue that if at the finest level the references are particles, once they start being very close to each other it may become impossible to keep them distinguishable, which makes it not possible to keep well established before/after relationship and therefore satisfy the alignment condition.
+One could argue that if both references and the object being measured are particles, their extent (i.e.
+the support of the wavefunction or of the field excitation) is comparable, which breaks the strictness condition.
+Alternatively, one can also argue that if we approach Planck scale, we cannot refine our references any longer, which breaks the refinement condition.
+The gist is that, in the case of quantities over which objects have an extent, the requirements for ordering can only be understood in terms of simplifying assumptions.
+
+The inability to create an ordering at the finest scale means we are not able to define other geometrical quantities, such as distances and angles.
+In this light, the ultimate structure of space-time may not be geometrical but it will still be topological, as we will still need to describe what is experimentally accessible.
+
+\subsection{Outline}
+
+The full mathematical account, which is available at [cite book], uses tools and constructions that, in our experience, are not widespread among physicists.
+Moreover, that level of detail may not be of interest to most.
+Therefore, we have recast the definitions and the main arguments to the most basic elements of set theory and order theory, which we believe are more accessible and cover the most important points.
+This should allow to give a more intelligible account of the physics without hiding it in the math.
+
+We will first review the link between point-set topology and experimental verifiability.
+We will show how a reference can be defined in terms of sets.
+We will give the basic notions of order theory we need.
+We will then formalize the requirements a set of references need to satisfy to identify continuous quantity.
+Finally, we will see how these requirements describe idealized conditions.
+
 \section{Elements of topology and its link to experimental verifiability}
@@
 The first thing we need to establish is the link between topology and experimental verifiability.
-Topology is, knowingly or unknowingly, widely used in physics as it is the foundation of many other tools, such as differential geometry and Lie groups.
+Point-set topology (or general topology) is, knowingly or unknowingly, widely used in physics as it is the foundation of many other tools, such as differential geometry and Lie groups.
 Most mathematical structures used in physics are topological spaces.
-Therefore it is crucial to understand what physical content do they capture.\footnote{In computer science, the link between topology and computability is already accepted though not widely known.
+Therefore it is crucial to understand what physical content they capture.\footnote{In computer science, the link between topology and computability is already accepted [TODO: add ref?] though not widely known.
 There is a link between those concepts and the ones presented here, but, for brevity, we are not going to expand on it.
@@
 If $X$ is the set of all possible cases, each verifiable statement is associated with a set $U \subseteq X$ and the statement can be re-expressed as ``$x$ is in $U$''.
-In our first example, $x$ is the mass of the neutrino and $U$ is the set of possible values from $0.045$ and $0.055$.
+In our first example, $x$ is the mass of the neutrino and $U$ is the set of possible values between $0.045$ and $0.055$.
 
@@
 In this case, as long as one statement is verified, the OR is verified and we can terminate.
-Therefore we do not care how many statements are there after the one that was verified, so the infinite union of verifiable sets is verifiable.
-Technically, topologies are closed under arbitrary unions, not just countable.
+Therefore we do not care how many statements there are after the one that was verified, so the infinite union of verifiable sets is verifiable.
+Note that topologies are closed under arbitrary unions, not just countable.
 Yet, because we cannot test more than countably many statements, the set of all verifiable statements must be constructable for a countable set of statements, or we would not be able to fully explore the space even with unlimited time.
@@
 These, in fact, correspond to all the finite precision measurements.
-It will also contains their unions.
-For example ``the absolute value of the charge of the electron is $1.6 \pm 0.005\;10^{-19}$ C'' would correspond to ``$x$ is in $(-1.65 \;10^{-19}, 1.55 \;10^{-19})\cup(1.55 \;10^{-19},1.65 \;10^{-19})$.
+It also contains their unions.
+For example ``the absolute value of the charge of the electron is $1.6 \pm 0.005\;10^{-19}$ C'' would correspond to ``$x$ is in $(-1.65 \;10^{-19}, -1.55 \;10^{-19})\cup(1.55 \;10^{-19},1.65 \;10^{-19})$.
 Singletons, sets with a single value, are not part of the topology and in fact we cannot verify them experimentally.
@@
 It captures the way that we can experimentally distinguish a set of possible cases.
-The main point is that statements have a binary logic in terms of TRUE/FALSE, while experimental tests have a ternary logic in terms of SUCCESS/FAILURE/UNDEFINED.
+The main point is that statements have a binary logic in terms of TRUE/FALSE, while experimental tests have a ternary logic in terms of SUCCESS/FAILURE/UNDEFINED, where undefined corresponds to non-termination.
 The topology is generally used to keep track of those differences.
@@
 What are the requirements on our experimental apparatus such that those quantities can be measured?
-Can we expect those requirements to hold when we go to ever smaller scale?
+Can we expect those requirements to hold when we go to ever smaller scales?
 If not, how do they break down?
 
-Since we established a link between topology and experimental verifiability, we can pose that questions in a way that is both mathematically and physically precise.
-Suppose we have a set of possible cases $X$, which we will call possibilities, and a way to distinguish them experimentally, which means we have a topology $\mathsf{T}$ that captures all our verifiable sets.
+Since we established a link between topology and experimental verifiability, we can pose the above questions in a way that is both mathematically and physically precise.
+Suppose we have a set of possible cases $X$, which we will call \textbf{possibilities}, and a way to distinguish them experimentally, which means we have a topology $\mathsf{T}$ that captures all our verifiable sets.
 What are the requirements on $(X, \mathsf{T})$ such that it is homeomorphic (i.e.
@@
 
-The first thing we need to develop is an abstract model to represent how quantities are measured experimentally. Let us first gather some requirements from the physics.
+The first thing we need to develop is a conceptual model to represent how quantities are measured experimentally. Let us first gather some requirements from the physics.
 
 We start from the notion of a reference, such as a mark on a ruler or the tick of a clock.
-A reference, then is a physical object that allows us to distinguish between three cases: a before, an on and an after the reference.
+A reference, then, is a physical object that allows us to distinguish between three cases: a before, an on and an after the reference.
 That is, a point can be before, on or after the mark on a ruler; an event can happen before, on or after the tick of a clock.
@@
 In general, the three cases a reference defines are not mutually exclusive: an object can extend before, on and after the mark; an event can start before and end after the tick of a clock.
-In fact, if the object is both before and after, we expect it to be on as well.
+In fact, if the object is both before and after, it will be on the reference as well.
 In all cases, the object should be found at least in one of the three cases.
 
-If a reference is at the end of the range, then it will not have an after or a before. Yet, a reference is a physical object and needs to take some space, therefore there will always be an on case.
+If a reference is at an end of the range, then it will either not have an after or a before.
+Yet, a reference is a physical object and needs to take some space, therefore there will always be an on case.
 
@@
 
-If $(X, \mathsf{T})$ is the set of physically distinguishable cases, we can formalize a reference as a triplet $r = (B, O, A)$ of three subsets $B, O, A \subseteq X$.
-Respectively, they will represent the cases in which the object is before, on or after the references.
+If $(X, \mathsf{T})$ is the topological space of the physically distinguishable cases, the number of ways an object can be found to be, we can formalize a reference as a triplet $r = (B, O, A)$ of three subsets $B, O, A \subseteq X$.
+Respectively, they will represent the cases in which the object is before, on or after the reference.
 We can capture the rest of the requirements with the following:
@@
 A measuring device will be composed of several references.
-For a continuous quantities, we can intuitively understand that we are assuming that, though we always have a finite set with finite precision, we assume we can refine our instruments with ever greater precision.
+Intuitively, for continuous quantities we are assuming that, though we always have a finite set of references with finite precision, we could in principle refine our instruments with ever greater precision.
 Our work is to make this intuitive notion precise.
@@
 There are many ways to mathematically characterize the real numbers.
-The one we are interested is in terms of their order, instead of operations like addition or multiplication, as it is the one that imposes the least requirements and it is enough to identify the topology.\footnote{Note how all transformations that preserve addition or multiplication must preserve the ordering, while the converse is not true.
-For example, non-linear monotonic transformations preserve the ordering and the topology while they do not preserve the group structure.} Let us review, then, a few key concepts of order theory, which is the branch of mathematics that studies ordered sets and their properties.
+The one we are interested in is in terms of their order, instead of operations like addition or multiplication, as it is the one that imposes the least requirements and it is enough to identify the topology.\footnote{Note how all transformations that preserve addition or multiplication must preserve the ordering, while the converse is not true.
+For example, consider $x \to x^3$, which is a non-linear monotonic transformation.
+It preserves the ordering (i.e.
+$x_1 \leq x_2$ if and only if $x_1^3 \leq x_2^3$) while it does not preserve addition (i.e.
+$(x_1 + x_2)^3 \neq x_1^3 + x_2^3$).} Let us review, then, a few key concepts of order theory, which is the branch of mathematics that studies ordered sets and their properties.
 
-A partial order $\leq : X \times X \to \mathbb{B}$ is a binary relationship that is reflexive (i.e.
+A partial order $\leq$ is a binary relationship that is reflexive (i.e.
 $a \leq a$), antisymmetric (i.e.
-if $a\leq b$ and $b \leq a$ then $a = a$) and transitive (i.e.
+if $a\leq b$ and $b \leq a$ then $a = b$) and transitive (i.e.
 if $a \leq b$ and $b \leq c$ then $a \leq c$).
+In a partial order, two distinct elements are not necessarily one before the other.
+On a plane, we can order points by their horizontal position, yet that would not order points that lie on a vertical line.
 A total order, or linear order, is a partial order such that any two elements are comparable (i.e.
 at least $a \leq b$ or $b \leq a$).
+If not explicitly stated, we will assume orders to be linear.
 
 Given an ordered set $(X, \leq)$ we can construct the order topology in the following way.
-Take all sets of the form $(a, \infty) = \{x \in X \, | \, a < x\} \;,\; (-\infty, b) = \{x \in X \, | \, x < b\}$.
-This is a basis for the topology.
-Then take all the sets that can be generated from the basis through finite intersection and arbitrary union.
+Take all sets of the form $(a, \infty) = \{x \in X \, | \, x > a\} \;,\; (-\infty, b) = \{x \in X \, | \, x < b\}$.
+This will be a basis for the order topology.
+We then take all the sets that can be generated from the basis through finite intersection and arbitrary union.
 This will be the order topology.
 Both the integers and the reals are totally ordered sets with their standard $\leq$ relationship.
-Their respective order topology correspond to their standard topology.
-Note how each set within the basis corresponds a statement like ``x is after a'' or ``x is before b''.
+Their respective order topologies correspond to their standard topologies.
+Note how each set within the basis corresponds to a statement like ``x is after a'' or ``x is before b''.
 Note how these, already, are very similar to the before and after cases we introduced in the previous section.
 
-Like groups or topological spaces, two ordered sets are isomorphic if their have equivalent structure.
-That is, two ordered sets are isomorphic if there is a bijection that preserves the ordering.
-If two ordered set are isomorphic, then they will have the same order topology and vice-versa.
+Like groups or topological spaces, two ordered sets are isomorphic if they have equivalent structure.
+That is, two ordered sets are isomorphic if there is a bijection that preserves the ordering (i.e.~an invertible monotonic function).
+If two ordered sets are isomorphic, then they will have the same order topology and vice-versa.
 So the ordering identifies the topology of an ordered set and vice-versa.
@@
 In light of this, we should break our problem into two parts.
-First understand what are the requirements on references such that the set of possibilities $X$ is totally ordered and its topology is the order topology.
-Then what are the additional requirements to be ordered like the real numbers or the integers.
+First understand what the requirements are on references such that the set of possibilities $X$ is linearly ordered and its topology is the order topology.
+Then understand what the additional requirements are to be ordered like the real numbers or the integers.
 It turns out that most of the work lies in the first part, in recovering the linear order.
 
-\section{Experimental requirements for construct real valued quantities}
+\section{Experimental requirements for constructing real-valued quantities}
 
@@
 That is, a reference $r= (B, O, A)$ is strict if $B \cap O = \emptyset$, $O \cap A = \emptyset$ and $B \cap A = \emptyset$.
-Technically, we would just need that before and after are disjoint ($B \cap A = \emptyset$) and then redefine on as what remains ($O = X / (A \cup B)$).
+Technically, we would just need that before and after are disjoint ($B \cap A = \emptyset$) and then redefine on as what remains ($O = X \setminus (A \cup B)$).
 That is, as long as before and after are separate cases, we can redefine on as ``not before and not after''.
 
-The order topology allows us to construct strict references.
+The order topology always allows us to construct strict references.
 If we take $a,b \in \mathbb{R}$ such that $b<a$, we have $( (-\infty, b), [b, a], (a, +\infty) )$ that represents a reference that physically extends from $b$ to $a$.
 Mathematically, we can see that the extent of the reference is represented by a falsifiable (i.e.
-closed).
-Conversely, if we want to be able to construct the order topology from references, then, those will need to be strict.
-Intuitively if we can experimentally define an order on all possibilities, then, given two distinct ones, we must be able to find references that are able to confirm unambiguously that one is before the other.
+closed) set $[b,a]$, while the others are verifiable (i.e.
+open) sets.
+Therefore if we want to rederive the order topology starting from references alone, those will need to be strict.
+Intuitively, defining an order experimentally on all possibilities means that, given two distinct ones, we are able to find references that will confirm unambiguously that one possibility is before the other.
 Therefore we must be able to tell before and after apart.
 
-What does requirement entails in practice?
-References for quantities that we count (i.e.
-discrete quantities) can always be made strict: either you have $n$ elements, less than $n$ or more than $n$.
+What does this requirement entail in practice?
+References for quantities that we count (i.e.~discrete quantities) can always be made strict: either you have $n$ elements, less than $n$ or more than $n$.
 The reason is that the quantity does not really have an extent and the possible values are well separated.
-Quantities over which we have an extent, like space and time, are another matter.
+[TODO: Clarify to avoid confusion] Quantities over which we have an extent, like space and time, are another matter.
 If the extent of the object being measured is much smaller than the extent of the reference, then the object cannot be found both before and after.
@@
 As we put references together, we must be sure they are related to each other in the correct way if we want to end up with a linear ordering.
-Intuitively, if we mix reference of horizontal position with vertical positions we will not end up with a linear ordering.
+Intuitively, if we mix references for horizontal position with references for vertical position we will not end up with a linear ordering.
 
-To define what it means for two references to be pointing in the same direction, take two values $b_1, b_2 \in \mathbb{R}$.
+To define what it means for two references to be aligned, take two values $b_1, b_2 \in \mathbb{R}$.
 Suppose that $b_1 \leq b_2$.
@@
 In terms of sets, $(-\infty, b_1) \subseteq (-\infty, b_2)$.
-In a linear order, the idea that one point is before the other can be translate in terms of set inclusion.
-Alignment between references, then, can be defined by requiring that the before and after statement have an inclusion relationship.
+In a linear order, the idea that one point is before the other can be translated in terms of set inclusion.
+Alignment between references, then, can be defined by requiring that the before and after statements have an inclusion relationship.
 
-We also note that the negation of ``$x$ is greater than $a$ is ``$x$ is less or equal to $a$''.
+We also note that the negation of ``$x$ is greater than $a$'' is ``$x$ is less than or equal to $a$''.
 In terms of sets, $(a, +\infty)^C = (-\infty, a]$.
-For a strict reference extending from $a$ to $b$, we will have $(-\infty, b) \subseteq (-\infty, a]$.
+For a strict reference extending from $b$ to $a$, we will have $(-\infty, b) \subseteq (-\infty, a]$.
 
@@
 If $B_1 \subset B_2 \subset A_1^C \subset A_2^C$ then the two references are overlapping.
-Alignment of references is another necessary condition if we want to reconstruct and ordering that is linear.
+Alignment of references is another necessary condition if we want to reconstruct an ordering that is linear.
 
-What does this requirement entails in practice?
+What does this requirement entail in practice?
 It means that there are clear fixed ordering relationships between the references.
@@
 That is, if $O_2 \subset O_1$ then we can find another reference such that $O_3 \subset O_1$ and $O_2 \cap O_3 = \emptyset$.
-If a set of references have this property, then we say it is refinable: the references can be refined to non overlapping references that cover all the possibilities.
+If a set of references has this property, then we say it is refinable: the references can be refined to non-overlapping references that cover all the possibilities.
 
-What does this requirement entails in practice? It means that we can place a reference wherever we want and can shrink their extent to the point they occupy only one possibility.
+What does this requirement entail in practice? It means that we can place a reference wherever we want and can shrink its extent such that it occupies only one possibility.
 
@@
 The idea is that, under those conditions, we can find references that are so fine that they extend over only one possibility.
-Therefore we have a one to one correspondence between experimentally distinguishable cases and the finest references.
+Therefore we have a one-to-one correspondence between experimentally distinguishable cases and the finest references.
 Since all references are aligned and strict, two different references must be sequential, one before the other.
@@
 
-The order of the integers can be characterized as the one that does not have a minimum, does not have a maximum and, taken any two elements, there are finitely many ones.
+The order of the integers can be characterized as the one that does not have a minimum, does not have a maximum, and given any two elements, there are finitely many elements between them.
 
@@
 
-The order of the reals can be characterized as one that is dense (i.e.
-between two elements there is always another one), complete (i.e.
-contains all the limit points), has a countable dense subset (i.e.
-between two real numbers we can always find a rational, which is a countable set), has no minimum or maximum.
+The order of the reals can be characterized as one that is dense (i.e.~between two elements there is always another one), complete (i.e.~it contains all the limit points), has a countable dense subset (i.e.~between two real numbers we can always find a rational, which is a countable set), and has no minimum or maximum.
 
@@
 The set $Q$ is dense in the original set $X$ and the original set $X$ is complete.
-Therefore the $X$ is the completion of $Q$.
-Another theorem in order theorem states that the completion of any order is unique.
-Since the real numbers are the completion of the rationals, and $Q$ has the same order of the rational, then $X$ has the same order of the reals.
+Therefore $X$ is the completion of $Q$.
+Another theorem in order theory states that the completion of any order is unique.
+Since the real numbers are the completion of the rationals, and $Q$ has the same order of the rationals, then $X$ has the same order of the reals.
 
@@
 The completion and the countable dense subset are a consequence of the more general topological requirements.
-The inclusion of all the limit points will come from the infinite union allowed by the topology while the dense subset is associated with the countable base, which is a requirement for out topology to be physically meaningful.
+The inclusion of all the limit points will come from the infinite union allowed by the topology while the dense subset is associated with the countable base, which is a requirement for our topology to be physically meaningful.
 
@@
 \begin{enumerate}
-	\item the ability to compare the quantity (i.e. position, time, mass, ...) with references such that we can say whether the object is before or after each one
+	\item the ability to compare the quantity (i.e.~position, time, mass, ...) with references such that we can say whether the object is before or after each one
 	\item the three different cases, before/on/after, must be mutually exclusive
@@
 Under these conditions, we can assign a real number to the finest possible references, which will correspond to all possible distinguishable cases.
-Higher dimension can be reconstructed by combining sets of references for different quantities, which leads to the familiar structure of a manifold.
+Higher dimensions can be constructed by combining sets of references for different quantities, which leads to the familiar structure of a manifold.
 
@@
 We are not, therefore, positing that there already exists a continuous quantity to be found.
-We are asking, under what conditions we can label the outcomes of our measurement scheme with a continuous quantity.
-This in line with the perspective of considering a physical theory, including quantum theory, as dealing with the output of measurements.
-For example, Wheeler's commented: ``It is wrong, moreover, to regard this or that physical quantity as sitting out therewith this or that numerical value''.
-What we are doing is go at a more fundamental level to understand this works in the continuous case, and how can it be modeled in a rigorous say.
+We are asking under what conditions we can label the outcomes of our measurement scheme with a continuous quantity.
+This is in line with the perspective of considering a physical theory, including quantum theory, as dealing with the output of measurements.
+For example, Wheeler commented: ``It is wrong, moreover, to regard this or that physical quantity as sitting out there with this or that numerical value''.[TODO: Add ref] What we are doing is going to a more fundamental level to understand how this works in the continuous case, and how can it be modeled in a rigorous say.
 
-Now that we understand what the requirements are we can ask: are these requirement tenable? Or are they necessarily idealizations?
+Now that we understand what the requirements are we can ask: are these requirements tenable? Or are they necessarily idealizations?
 
@@
 That is, the failure of any one of the first four means that ordered quantities are not physically meaningful.
-Which means we cannot substitute the real number with the rationals or the integers, as these are themselves ordered.
-Note that all geometrical structure are based on some notion of numeric distance, therefore if order fails geometrical structure fail as well.
+This means we cannot substitute the real numbers with the rationals or the integers, as these are themselves ordered.
+Note that all geometrical structures are based on some notion of numeric distance, therefore if order fails geometrical structures fail as well.
 We are not going to be able to define metric tensors, symplectic forms or even a differentiable structure.
 
-As such, the bar to show that, at a fine level, space-time does not posses a geometrical structure is very low.
+As such, the bar to show that, at a fine level, space-time does not possess a geometrical structure is very low.
 One simply has to instill the doubt that any of (i-iv) is untenable.
@@
 For (ii), we saw that references are objects that have, in general, an extent in space.
-A reference can be considered strict if the extent if the extent of the object measured is much smaller than the extent of the reference.
+A reference can be considered strict if the extent of the object measured is much smaller than the extent of the reference [TODO: Fix however gets fixed above].
 For example, the top level of a liquid in a jar is thinner than the marks used to measure the content.
@@
 Therefore, if we assume that we can make finer and finer references, at some point we will find the extent of the references comparable to the extent of the objects we are measuring.
-This is also true for fundamental particles, where the extent of the system correspond to the extent of the wave function: if the wavefunctions overlap, one particle is not clearly before/on/after the other.
+This is also true for fundamental particles, where the extent of the system corresponds to the extent of the wave function: if the wavefunctions overlap, one particle is not clearly before/on/after the other.
 Similar arguments apply for excitations of quantum fields.
@@
 
-For (iv), the requirement for refinement means we can always prepare finer and finer references. This is generally thought to not be possible once length scales are comparable to the Planck length.
+For (iv), the requirement for refinement means we can always prepare finer and finer references. This is generally thought not to be possible once length scales are comparable to the Planck length.
 
 We are not interested here to expand on these arguments or pick ones that are universally accepted.
-That will be the scope for future work.
-Here we simply want to show that is relatively easy to formulate reasonable arguments.
-As long as one finds reasonable that one of (i-iv) fails, then our point follows: space-time structure at a fine scale will not be geometrical, will not be ordered, but it will yet be topological, we will have some way to measure what can physically be well defined.
+That will be the scope of future work.
+Here we simply want to show that it is relatively easy to formulate reasonable arguments.
+As long as one finds reasonable that one of (i-iv) fails, then our point follows: space-time structure at a fine scale will not be geometrical, will not be ordered, but it will still be topological, giving us some way to measure what can physically be well defined.[TODO: Check that this last sentence means what we want it to.
+Maybe clarify what types of properties can still be physically well defined and distinguished/measured at these scales.
+Avoid the word measure, as it's usually associated with quantities?
+Would a better word be distinguish?]
 
@@
 
-It is not clear what topological space should be used instead of a manifold. There are
+We have seen what are the requirements to give rise to the real numbers through a set of of experimentally verifiable statements, which captures the most basic elements of how continuous quantities are measured in practice.
+We have also seen that those can only be considered idealized conditions, and that, when those requirements fail, ordering itself fails and no geometrical structure can be constructured.
+Naturally, this opens the question of what topological spaces would be appropriate in those regimes.
+This is something we do not have an answer for, though we can mentioned two of the possibly many scenarios.
 
-The simplest solution would be to find a topological space
-
-We leave
-
-Naturally, this opens the question of what topological spaces would be appropriate in those regimes.
 One can imagine to try and construct a topological space that is not ordered at a fine scale but is ordered at a large scale.
 A difficulty here is that typically the notion of distance is absent from topological spaces.
+Though it is not clear to us whether this is possible, at least it is a clear venue to explore.
 
-There is, however, a much more problematic prospect.
-General relativity tells us that the geometry depends on the content of the space.
-What is in the space affects the notion of distance.
-It would not be far fetch to assume that what is in the space affects what can be distinguished experimentally.
-This would mean that the topology itself would also depend on the content of the space.
-In that case, we would not have a fixed topology, in the same way that we do not have a fixed geometry in general relativity.
-Conceptually, we would need to define some sort of topological equation.
-Though the premise seems reasonable enough, it is not clear to us whether appropriate mathematical tools for such a line of inquiry currently exist.
-
-\section{Conclusion}
+There is, however, a much more drastic scenario.
+General relativity tells us that the geometry depends on the content of the space (i.e.
+the energy-momentum tensor).
+What is in the space affects the notion of distance and angle.
+It would not be far fetched to assume that what is in the space also affects what can be distinguished experimentally.
+As the energy-matter distribution changes the geometry by curving space-time, it would, at a more fundamental level, change the topology as well.
+It is not clear to us how one would even start exploring such scenario.
 

### 2020-02-14 Articles/2019-QuantumComposition/Source/QuantumComposition.tex
Revert "Revert "First pass at ICNFP proceeding""
This reverts commit 10675ec6434351e6f656673d7b6200762af56df5.

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}

\title{Quantum system composition}
\author{gabriele.carcassi }
\date{October 2019}

\begin{document}

\maketitle

\section{First try}

Postulate. The state of the system is represented by a unit vector $a$ in a vector space $\mathcal{A}$ with inner product $< \cdot, \cdot >_\mathcal{A} : \mathcal{A} \times \mathcal{A} \to \mathbb{C}$.

%Postulate 2. For each measurable quantity $q$ of the system there exists an Hermitian operator $O_q : \mathcal{A} \to \mathcal{A}$ such that the expectation value is given by $\bar{q}=<v, O_qv>_\mathcal{A}$.

Definition 1. A composite system $\mathcal{C}$ of two subsystems $\mathcal{A}$ and $\mathcal{B}$ is such that:
\begin{enumerate}
    \item given the states of the two subsystems we can construct a state in which those subsystems have that state; that is, there exists an injective map $C : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$

    \item when the state of one of the subsystems is fixed, the state space of the composite system is equivalent to the one of the free subsystem. That is, given a unitary vector $b \in \mathcal{B}$ then for all $a_1, a_2 \in \mathcal{A}$:

    $$<a_1, a_2>_\mathcal{A} = <C(a_1, b), C(a_2, b)>_\mathcal{C}.$$

    Similarly, given a unitary vector $a \in \mathcal{A}$ then for all $b_1, b_2 \in \mathcal{B}$:

    $$<b_1, b_2>_\mathcal{B} = <C(a, b_1), C(a, b_1)>_\mathcal{C}.$$

%\item every measurable quantity on each subsystem corresponds to a measurable quantity on the composite system with the same expectation value; that is, for every Hermitian operator $O_{q_a}^\mathcal{A} : \mathcal{A} \to \mathcal{A}$ associated with a quantity $q_a$ there exists an Hermitian operator $O_{q_a}^\mathcal{C} :\mathcal{C} \to \mathcal{C}$ such that
%$$\bar{q_a}=<v_a, O_{q_a}^\mathcal{A}v_a>_\mathcal{A} = <C(v_a, v_b), O_{q_a}^\mathcal{C}C(v_a, v_b)>_\mathcal{C}$$
%for all $v_a \in V_\mathcal{A}$, $v_b \in V_\mathcal{B}$. Similarly, for every Hermitian operator $O_{q_b}^\mathcal{B} : \mathcal{B} \to \mathcal{B}$ associated with a quantity $q_b$ there exists an Hermitian operator $O_{q_b}^\mathcal{C} :\mathcal{C} \to \mathcal{C}$ such that
%$$\bar{q_b}=<v_b, O_{q_b}^\mathcal{B}v_b>_\mathcal{B} = <C(v_a, v_b), O_{q_a}^\mathcal{C}C(v_a, v_b)>_\mathcal{C}$$
%for all $v_a \in V_\mathcal{A}$, $v_b \in V_\mathcal{B}$.
\end{enumerate}

Proposition 1. $C = \otimes $

Proof. Step 1. $C$ is bilinear. Fixing a unitary $b \in \mathcal{B}$, for all $a, a_1, a_2 \in \mathcal{A}$ and $k_1, k_2 \in \mathbb{C}$ we have
\begin{equation}
	\begin{aligned}
	<C(a, b), C(k_1a_1 + k_2a_2, b)>_\mathcal{C} &= <a, k_1a_1 + k_2a_2>_\mathcal{A} \\
	&= k_1<a, a_1>_\mathcal{A} + k_2<a, a_2>_\mathcal{A} \\
	&= k_1<C(a, b), C(a_1, b)>_\mathcal{C} + k_2<C(a, b), C(a_2, b)>_\mathcal{C} \\
	&= <C(a, b), k_1C(a_1, b) + k_2C(a_2, b)>_\mathcal{C} \\
	\end{aligned}
\end{equation}
Since this is valid for all $a, a_1, a_2 \in \mathcal{A}$ and $k_1, k_2 \in \mathbb{C}$ then $C(k_1a_1 + k_2a_2, b)=k_1C(a_1, b) + k_2C(a_2, b)$. Similarly, we have $C(a, k_1b_1 + k_2b_2)=k_1C(a, b_1) + k_2C(a, b_2)$. $C$ is bilinear.

%Step 1: inner product agrees over all linear operators. Let $A^\mathcal{A} : \mathcal{A} \to \mathcal{A}$ be a linear operator. We can decompose $A^\mathcal{A} = A_a^\mathcal{A} + \imath A_b^\mathcal{A}$ such that $A_a^\mathcal{A}$ and $A_b^\mathcal{A}$ are Hermitian. For all $v_a \in V_\mathcal{A}$, $v_b \in V_\mathcal{B}$ we have:
%\begin{equation}
%	\begin{aligned}
%	<v_a, A^\mathcal{A}v_a>_\mathcal{A} &= <v_a, A_a^\mathcal{A}v_a>_\mathcal{A} + \imath <v_a, A_b^\mathcal{A}v_a>_\mathcal{A} \\
%	&= <C(v_a, v_b), A_a^\mathcal{C}C(v_a, v_b)>_\mathcal{C} + \imath <C(v_a, v_b), A_b^\mathcal{C}C(v_a, v_b)>_\mathcal{C} \\
%	&= <C(v_a, v_b), A^\mathcal{C}C(v_a, v_b)>_\mathcal{C}
%	\end{aligned}
%\end{equation}
%Similarly for $\mathcal{B}$.

Step 2: C is the tensor product. As $C : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a bilinear operator, for the universal property of the tensor product we can find a linear operator $\hat{C} : \mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that $C(a, b) = \hat{C}(a \otimes b)$. The set $E = \{ a\otimes b | <a,a>=1, <b,b>=1 \}$ forms a basis for the tensor product and therefore $\hat{C}(E)$ is a basis for $\mathcal{C}$. We have :
$$<C(a, b), C(a, b)>_\mathcal{C} = <a\otimes b, a \otimes b>_{\otimes} = <a, a>_\mathcal{A} = <b, b>_\mathcal{B} = 1$$
The function $\hat{C}$, then, is unitary over all elements of the basis and is therefore the identity. We have $C(a, b) = \hat{C}(a \otimes b) = a \otimes b$.

\section{Second try}

Lemma 1. Let $\mathcal{A}$, $\mathcal{B}$ and $\mathcal{C}$ be linear spaces. Let $C : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ be an injective map that satisfy the following:
\begin{enumerate}
	\item given a positive definite linear operator $\rho_\mathcal{A} : \mathcal{A} \to \mathcal{A}$, there exists a corresponding positive definite linear operator $\rho_\mathcal{C} : \mathcal{C} \to \mathcal{C}$ such that $\rho_\mathcal{C}(C(a,b)) = C(\rho_\mathcal{A}(a),b)$; moreover $k_1\rho_\mathcal{A}^1 + k_2\rho_\mathcal{A}^2$ corresponds to $k_1\rho_\mathcal{C}^1 + k_2\rho_\mathcal{C}^2$
	\item given a positive definite linear operator $\rho_\mathcal{B} : \mathcal{B} \to \mathcal{B}$, there exists a corresponding positive definite linear operator $\rho_\mathcal{C} : \mathcal{C} \to \mathcal{C}$ such that $\rho_\mathcal{C}(C(a,b)) = C(a,\rho_\mathcal{B}(b))$; moreover $k_1\rho_\mathcal{B}^1 + k_2\rho_\mathcal{B}^2$ corresponds to $k_1\rho_\mathcal{C}^1 + k_2\rho_\mathcal{C}^2$
\end{enumerate}
Then $C$ is a bilinear operator. That is, $C(k_1a_1 + k_2a_2, b)=k_1C(a_1, b) + k_2C(a_2, b)$ and $C(a, k_1b_1 + k_2b_2)=k_1C(a, b_1) + k_2C(a, b_2)$.

Proof. \emph{Extend linearity to negative and imaginary operators.} We can univocally extend the correspondence to negative and imaginary operators. If $\rho_\mathcal{A}$ is a positive operator, we have $\rho_\mathcal{A} = - (-\rho_\mathcal{A})$. If we extend the linear correspondence, we must have $\rho_\mathcal{C} = - (-\rho_\mathcal{A})_\mathcal{C}$ and therefore $(-\rho_\mathcal{A})_\mathcal{C} = - \rho_\mathcal{C}$. Similarly, $- \rho_\mathcal{A} = \imath (\imath \rho_\mathcal{A})$ which leads to $(\imath \rho_\mathcal{A})_\mathcal{C} = \imath \rho_\mathcal{C}$. One can perform the same extension on $\mathcal{B}$.

\emph{All linear operators have a linear correspondence.} Let $O_\mathcal{A}$ be a linear operator. We can always write $O_\mathcal{A} = O_\mathcal{A}^{r+} - O_\mathcal{A}^{r-} + \imath O_\mathcal{A}^{i+} - \imath O_\mathcal{A}^{i-}$ where  $O_\mathcal{A}^{r+}$, $O_\mathcal{A}^{r-}$, $\imath O_\mathcal{A}^{i+}$ and $\imath O_\mathcal{A}^{i-}$ are all positive definite. Therefore we can find a corresponding $O_\mathcal{C}$. One can perform the same expansion on linear operators on $\mathcal{B}$.

\emph{The map $C$ is bilinear.} Let $a_1, a_2 \in \mathcal{A}$. The we can find a linear operator such that $a_2 = O a_1$. We have $C(k_1 a_1 + k_2 a_2, b) = C(k_1 I a_1 + k_2 O a_1, b) = (k_1 I + k_2 O)_\mathcal{C}C(a_1, b) = (k_1 I)_\mathcal{C}C(a_1, b) + (k_2 O)_\mathcal{C}C(a_1, b) = C(k_1 I a_1, b) + C(k_2 O a_1, b)  = C(k_1 a, b) + C(k_2 a_2, b)$. Similarly, we find $C(a, k_1b_1 + k_2b_2)=k_1C(a, b_1) + k_2C(a, b_2)$. The map is bilinear.

\end{document}

### 2020-02-14 Proceedings/2019-ICNFP/Carcassi2019ICNFPProceedings.tex
Revised first section

@@
 In our project, Assumptions of Physics, the idea is to find a minimal set of assumptions from which the different basic theories can be derived.
-The idea is that, by doing this as formally as possible, we are forced to specify all our implicit starting points and inform us to what can be done different.
+The idea is that, by doing this as formally as possible, we are forced to specify all our implicit starting points, thus clarifying what could possibly be done differently.
 While other approaches start with the same premise, we typically start at a much lower level.
-For example, we do not assume probability spaces which are already constructions upon sigma-algebras and measures, and sigma-algebras are typically constructed from the topology of the space.
-We aim to recover those as well.
-This is particular important if one is interested in the structure of space-time, since its topology and sigma-algebra are the most foundational aspects.
+For example, we do not assume probability spaces, which are already constructions upon sigma-algebras and measures, and sigma-algebras are typically constructed from the topology of the space.
+We instead aim to recover those as well.
+This is particularly important if one is interested in the structure of space-time, since its topology and sigma-algebra are the most foundational aspects.
 
-The other difference is that our approach wants to start with primitives that are necessary to do physics.
+The other difference is that our approach aims to start with primitives that are necessary to do physics.
 Our basic building block is the notion of a verifiable statement: an assertion for which an experimental test is available that would confirm, in finite time, that the assertion is true.
@@
 That is, in physics, the topology captures what is experimentally verifiable (i.e.
-$U$ is an open set if and only if ``$x$ is $U$'' is experimentally verifiable).
-For real valued quantities, the verifiable statement correspond to open intervals to signify that our measurement have always finite precision.
-The fact that a continuous quantity has precisely a given real value (e.g.
-the length of the side is exactly 1 meter and the diagonal is exactly $\sqrt{2}$ meters) is not experimentally verifiable.
+$U$ is an open set if and only if ``$x$ is in $U$'' is experimentally verifiable).
+For real valued quantities, the verifiable statements correspond to open intervals to signify that our measurements always have finite precision.
+The fact that a continuous quantity has precisely a given real value (e.g.~the length of the side is exactly 1 meter and the diagonal is exactly $\sqrt{2}$ meters) is not experimentally verifiable.
 This gives new insight to the notion of topological continuity and it tells why, in very general terms, functions \textbf{must} be well behaved in physics.
@@
 This idea also clarifies the division between topology and geometry.
-Topological construct are more primitive than geometrical ones because they have no notion of size.
+Topological constructs are more primitive than geometrical ones because they have no notion of size.
 Physically, it means that before we can assign distances between two points, we must be able to distinguish them, to tell them apart.
@@
 The geometry is Riemmanian.
-On phase space, within a degree of freedom, units of position cannot be compared to unit of momentums, therefore we cannot define angles.
-Yet, areas of a degree of freedom are proportional to the number of possible configurations.
+On phase space, within a degree of freedom, units of position cannot be compared to unit of momentum, therefore we cannot define angles.
+Yet, phase-space areas for a degree of freedom are proportional to the number of possible configurations.
 The fact that we can quantify areas for each degree of freedom is what gives us symplectic geometry.
-If we imagine the space of all possible blood work results, however, we have neither notions, and therefore we simply have a manifold that is not geometric.
+If we imagine for example the space of all possible blood work results, however, we have neither notion, and therefore we simply have a manifold that is not geometric.
 
@@
 Before we want to understand whether space-time is geometrically truly a Riemmanian manifold, we need to understand whether space-time is topologically truly a manifold.
-A manifold, simply put, is a set of possible cases $X$ that can be experimentally identified by a set of real valued.
+A manifold, simply put, is a set of possible cases $X$ that can be experimentally identified by a set of real values[TODO:ck that "real values" is what we want].
 So, the question is: when can a set of possible cases, or possibilities for short, be experimentally identified by a real value?
@@
 
-Even though this work is mathematically very technical and abstract, we strive to have a well understood map to more tangible and physical concepts: that is one of the main goals in our work.
+Even though this work is mathematically very technical and abstract, we strive to have a well understood map to more tangible and physical concepts; that is one of the main goals in our work.
 Therefore let us first give a summary of the results in physics terms, leaving the details to subsequent sections.
@@
 Mathematics offers many constructions of the real numbers from more primitive notions.
-Common techniques include completion through Cauchy sequences, using Dedekind cuts or adding constraints over an algebraic fields.
+Common techniques include completion through Cauchy sequences, using Dedekind cuts or adding constraints over an algebraic fields[TODO:ck singular or plural]].
 In fact, Faltin et al.
-(cite) says:
+(TODO:cite) states:
 \begin{quote}
@@
 A single reference is something that allows us to distinguish a before and an after, a smaller and a greater, a higher and a lower.
-We can formalize a reference with our more basic notion of a verifiable statement: a reference gives us two verifiable statements ``the objects is before this reference'' and ``the object is after this reference''.
-This allows to recast ordering relationship in terms of logical relationship for example, if reference A is after reference B, then if ``the object is before reference B'' is true, ``the object is before reference A'' must also be true.
-And these logical relationships remain conceptually the same regardless of whether ``before'' means ``before in time'' or ``before along a spatial direction''.
+We can formalize a reference with our more basic notion of a verifiable statement: a reference gives us two verifiable statements, ``the objects is before this reference'' and ``the object is after this reference''.
+This allows us to recast ordering relationship in terms of logical relationship.
+For example, if reference A is after reference B, then if ``the object is before reference B'' is true, ``the object is before reference A'' must also be true.
+And these logical relationships remain conceptually the same regardless of whether ``before'' means ``before in time'' or ``before along a spatial direction'', etc.
 These are the sort of basic ideas and language we use within our framework, and from these we work out a set of necessary and sufficient conditions that a set of references needs to satisfy so that they define a set of possible distinguishable cases, or possibilities for short, that is ordered like the real line.
@@
 If we are to measure the position of an object, then, it may extend before, on and after the mark of a ruler.
-In these cases we typically note the position of the beginning of the object and an end of the object.
+In these cases we typically note the position of the beginning of the object and an [TODO: an or the?] end of the object.
 This implicitly relies on the fact that the resolution of our eyes is higher than that of the ruler, and that we can independently recognize parts of the object, namely the beginning and end.
-Identifying parts over a higher resolution correspond, in the end, to new references.
+Identifying parts using a higher resolution corresponds, in the end, to new references.
 So, assuming this can always happen means implicitly assuming what we are tasked to derive.
 
-Unfortunately, thinking visually and drawing pictures is inherently misleading because everything we draw is inherently ordered.
-It confuses more than it helps, so much so that we progressed only when we stopped trying to reason with pictures and concentrated on just the boolean logic of our statements in truth tables.
+Unfortunately, thinking visually and drawing pictures is inherently misleading because everything we draw is inherently ordered[TODO: use inherently twice?].
+It confuses more than it helps, so much so that we ourselves progressed only when we stopped trying to reason with pictures and concentrated on just the boolean logic of our statements in truth tables.
 In that setting, the only elements we are allowed to define are those simple logical relationships we described before: if the object is before A, then it is not after B and it is before C.
@@
 The first major finding is that \emph{most of the conditions are not required by the real numbers specifically, but they are required by ordered quantities in general}.
-That is, the set of reference have to satisfy a few conditions just to be sure that the cases they distinguish experimentally are ordered (i.e.
-can be thought to be all being one after the other).
+That is, the set of reference has to satisfy a few conditions just to be sure that the cases they distinguish experimentally are ordered (i.e.~can be thought to be all being one after the other).
 The conditions are the following.
 \begin{itemize}
-	\item The references must be strict, meaning an object is either exclusively before, on or after the reference.
+	\item The references must be strict [TODO: Isn't this after they've already been sufficiently refined?], meaning an object is either exclusively before, on or after the reference.
 No overlaps are allowed.
@@
 	\item The references must be aligned, meaning their before and after statements identify incremental regions.
-In practice, this means we can always distinguish between them and they remain with constant before/after relationships (e.g.
-they do not fluctuate).
+In practice, this means we can always distinguish between them and they remain with constant before/after relationships (e.g.~they do not fluctuate).
 	\item The references must be refinable, meaning we can always resolve overlaps and fill in the whole space.
-For example, if an object can be between two references, then we must be able to put a reference as well.
+For example, if an object can be between two references, then we must be able to put a new reference between those two references as well.
 If we have two references that overlap, then we must be able to find other references that fit within that overlap, to tell what is before and after.
@@
 
-The second finding is that \emph{the inability to distinguish below a certain scale most likely invalidates those conditions, leading to a breakdown in ordering.} In other words, it is not just that the points of our space can't be labeled by real numbers: they cannot be ordered and therefore cannot be labeled by any ordered quantity.
+The second finding is that \emph{the inability to distinguish below a certain scale most likely invalidates those conditions, leading to a breakdown in ordering.} In other words, it is not just that the points of our space cannot be labeled by real numbers: they cannot be ordered and therefore cannot be labeled by any ordered quantity.
 To obtain that result, one must simply argue that at least one of the three conditions fails.
-For example, one could argue that if at the finest level the references are particles, once they start being very close to each other it may become impossible to keep them distinguishable, which makes it not possible to keep well established before/after relationship and therefore satisfy the alignment condition.
-One could argue that if both references and the object being measured are particles, their extent (i.e.
-the support of the wavefunction or of the field excitation) is comparable, which breaks the strictness condition.
+For example, one could argue that if at the finest level the references are particles, once they start being very close to each other it may become impossible to keep them distinguishable, which makes it impossible to keep well established before/after relationships and therefore satisfy the alignment condition.
+One could argue that if both references and the object being measured are particles, their extent (i.e.~the support of the wavefunction or of the field excitation) is comparable, which breaks the strictness condition.
 Alternatively, one can also argue that if we approach Planck scale, we cannot refine our references any longer, which breaks the refinement condition.
@@
 
-The full mathematical account, which is available at [cite book], uses tools and constructions that, in our experience, are not widespread among physicists.
+The full mathematical account, which is available at [TODO:cite book], uses tools and constructions that, in our experience, are not widespread among physicists.
 Moreover, that level of detail may not be of interest to most.
 Therefore, we have recast the definitions and the main arguments to the most basic elements of set theory and order theory, which we believe are more accessible and cover the most important points.
-This should allow to give a more intelligible account of the physics without hiding it in the math.
+This should allow us to give a more intelligible account of the physics without hiding it in the math.
 
@@
 We will give the basic notions of order theory we need.
-We will then formalize the requirements a set of references need to satisfy to identify continuous quantity.
+We will then formalize the requirements a set of references needs to satisfy to identify a continuous quantity.
 Finally, we will see how these requirements describe idealized conditions.

### 2020-02-21 Proceedings/2019-ICNFP/Carcassi2019ICNFPProceedings.tex
Update Carcassi2019ICNFPProceedings.tex

@@
 The geometry is Riemmanian.
-On phase space, within a degree of freedom, units of position cannot be compared to unit of momentum, therefore we cannot define angles.
+On phase space, within a degree of freedom, units of position cannot be compared to units of momentum, therefore we cannot define angles.
 Yet, phase-space areas for a degree of freedom are proportional to the number of possible configurations.
@@
 Before we want to understand whether space-time is geometrically truly a Riemmanian manifold, we need to understand whether space-time is topologically truly a manifold.
-A manifold, simply put, is a set of possible cases $X$ that can be experimentally identified by a set of real values[TODO:ck that "real values" is what we want].
+A manifold, simply put, is a set of possible cases $X$ that can be experimentally identified by a set of real values.
 So, the question is: when can a set of possible cases, or possibilities for short, be experimentally identified by a real value?
@@
 Mathematics offers many constructions of the real numbers from more primitive notions.
-Common techniques include completion through Cauchy sequences, using Dedekind cuts or adding constraints over an algebraic fields[TODO:ck singular or plural]].
+Common techniques include completion through Cauchy sequences, using Dedekind cuts or adding constraints over an algebraic field.
 In fact, Faltin et al.
@@
 A single reference is something that allows us to distinguish a before and an after, a smaller and a greater, a higher and a lower.
-We can formalize a reference with our more basic notion of a verifiable statement: a reference gives us two verifiable statements, ``the objects is before this reference'' and ``the object is after this reference''.
+We can formalize a reference with our more basic notion of a verifiable statement: a reference gives us two verifiable statements, ``the object is before this reference'' and ``the object is after this reference''.
 This allows us to recast ordering relationship in terms of logical relationship.
@@
 If we are to measure the position of an object, then, it may extend before, on and after the mark of a ruler.
-In these cases we typically note the position of the beginning of the object and an [TODO: an or the?] end of the object.
+In these cases we typically note the position of the beginning of the object and of the end of the object.
 This implicitly relies on the fact that the resolution of our eyes is higher than that of the ruler, and that we can independently recognize parts of the object, namely the beginning and end.
@@
 
-Unfortunately, thinking visually and drawing pictures is inherently misleading because everything we draw is inherently ordered[TODO: use inherently twice?].
+Unfortunately, thinking visually and drawing pictures is misleading because everything we draw is inherently ordered.
 It confuses more than it helps, so much so that we ourselves progressed only when we stopped trying to reason with pictures and concentrated on just the boolean logic of our statements in truth tables.
@@
 \begin{itemize}
-	\item The references must be strict [TODO: Isn't this after they've already been sufficiently refined?], meaning an object is either exclusively before, on or after the reference.
-No overlaps are allowed.
-In practice, it means that the extent of the object being measured is always much smaller than the extent of the reference, therefore it will always be considered wholly before, on or after the reference.
+	\item The references must be strict, meaning that the cases before, on and after are mutually exclusive.
+In practice, if objects have an extent, this must be much smaller than the extent of the reference, such that it will always be considered wholly before, on or after the reference.
 	\item The references must be aligned, meaning their before and after statements identify incremental regions.
@@
 There is a link between those concepts and the ones presented here, but, for brevity, we are not going to expand on it.
-The gist is that any computation device is also a physical system, and the output of a computation can be experimentally verified (i.e.
-we can read it).}
+The gist is that any computation device is also a physical system, and the output of a computation can be experimentally verified (i.e.~we can read it).}
 
@@
 Here we simply want to show that it is relatively easy to formulate reasonable arguments.
-As long as one finds reasonable that one of (i-iv) fails, then our point follows: space-time structure at a fine scale will not be geometrical, will not be ordered, but it will still be topological, giving us some way to measure what can physically be well defined.[TODO: Check that this last sentence means what we want it to.
-Maybe clarify what types of properties can still be physically well defined and distinguished/measured at these scales.
-Avoid the word measure, as it's usually associated with quantities?
-Would a better word be distinguish?]
+As long as one finds reasonable that one of (i-iv) fails, then our point follows: space-time structure at a fine scale will not be geometrical, will not be ordered, but it will still be topological, giving us some way to experimentally distinguish what can physically be well defined.
 
@@
 
-We have seen what are the requirements to give rise to the real numbers through a set of of experimentally verifiable statements, which captures the most basic elements of how continuous quantities are measured in practice.
-We have also seen that those can only be considered idealized conditions, and that, when those requirements fail, ordering itself fails and no geometrical structure can be constructured.
+We have seen what the requirements are to give rise to the real numbers through a set of experimentally verifiable statements, which captures the most basic elements of how continuous quantities are measured in practice.
+We have also seen that those can only be considered idealized conditions, and that, when those requirements fail, ordering itself fails and no geometrical structure can be constructed.
 Naturally, this opens the question of what topological spaces would be appropriate in those regimes.
-This is something we do not have an answer for, though we can mentioned two of the possibly many scenarios.
+This is something we do not have an answer for, though we can mention two of the possibly many scenarios.
 
@@
 There is, however, a much more drastic scenario.
-General relativity tells us that the geometry depends on the content of the space (i.e.
-the energy-momentum tensor).
+General relativity tells us that the geometry depends on the content of the space (i.e.~the energy-momentum tensor).
 What is in the space affects the notion of distance and angle.
-It would not be far fetched to assume that what is in the space also affects what can be distinguished experimentally.
+It would not be far-fetched to assume that what is in the space also affects what can be distinguished experimentally.
 As the energy-matter distribution changes the geometry by curving space-time, it would, at a more fundamental level, change the topology as well.
-It is not clear to us how one would even start exploring such scenario.
+It is not clear to us at this time how one would even start exploring such a scenario.
 
@@
 
+Funding for this work was provided in part by the MCubed program of the University of Michigan.
+
 \end{document}

### 2020-02-22 Proceedings/2019-ICNFP/Carcassi2019ICNFPProceedings.tex
Ready for submission

@@
 \section{Introduction}
+In our ongoing project, Assumptions of Physics, the idea is to find a minimal set of assumptions from which the different basic theories can be derived.
+The idea is that, by doing this as formally as possible, we are forced to specify all our implicit starting points, thus clarifying what could possibly be done differently.
+While other approaches start with a similar goal, see for example Refs.~\cite{PhysRevA.84.012311,QLogicReview,Hardy:2001jk,ludwig_hein_2013}, there are a number of key differences.
+One difference is that in our work both classical and quantum cases are derived on equal footing, identifying the key point of divergence between the two theories.\cite{Carc1} Another difference is that our approach aims to start with primitives that are necessary to do physics.
+Our basic building block is the notion of a verifiable statement: an assertion for which an experimental test is available that would confirm, in finite time, that the assertion is true.
+As one can imagine, without such a notion, we would have no experimental verification, which means no possibility of doing science.
+In general, the goal is to start with a tight connection between the physical concepts and their respective mathematical representations, so that it is always clear what the mathematical structure represents physically and no mathematical objects are unphysical.
 
-Find a few papers to cite that wonder about the structure of space time.
-
-In our project, Assumptions of Physics, the idea is to find a minimal set of assumptions from which the different basic theories can be derived.
-The idea is that, by doing this as formally as possible, we are forced to specify all our implicit starting points, thus clarifying what could possibly be done differently.
-While other approaches start with the same premise, we typically start at a much lower level.
+Another important difference is that we typically start at a lower level.
 For example, we do not assume probability spaces, which are already constructions upon sigma-algebras and measures, and sigma-algebras are typically constructed from the topology of the space.
@@
 
-The other difference is that our approach aims to start with primitives that are necessary to do physics.
-Our basic building block is the notion of a verifiable statement: an assertion for which an experimental test is available that would confirm, in finite time, that the assertion is true.
-As one can imagine, without such a notion, we would have no experimental verification, which means no possibility of doing science.
-
-In previous work, we have shown the link with this idea to the mathematical framework of topology.
+In previous work\cite{Carc2}, we have shown the link with this idea to the mathematical framework of topology.
 That is, in physics, the topology captures what is experimentally verifiable (i.e.
@@
 
-% We can cite others from wikipedi
-%F. Faltin, N. Metropolis, B. Ross and G.-C. Rota. The real numbers as a wreath product Advances in Mathematics, 16(1975), 278–304.
 Mathematics offers many constructions of the real numbers from more primitive notions.
@@
 In fact, Faltin et al.
-(TODO:cite) states:
+state\cite{faltin1975real}:
 \begin{quote}
@@
 
-The full mathematical account, which is available at [TODO:cite book], uses tools and constructions that, in our experience, are not widespread among physicists.
+The full mathematical account, which is available at \cite{Carc3}, uses tools and constructions that, in our experience, are not widespread among physicists.
 Moreover, that level of detail may not be of interest to most.
@@
 Most mathematical structures used in physics are topological spaces.
-Therefore it is crucial to understand what physical content they capture.\footnote{In computer science, the link between topology and computability is already accepted [TODO: add ref?] though not widely known.
+Therefore it is crucial to understand what physical content they capture.\footnote{In computer science, the link between topology and computability is already accepted, see e.g.
+\cite{escardo2004synthetic}, though not widely known.
 There is a link between those concepts and the ones presented here, but, for brevity, we are not going to expand on it.
@@
 The reason is that the quantity does not really have an extent and the possible values are well separated.
-[TODO: Clarify to avoid confusion] Quantities over which we have an extent, like space and time, are another matter.
-If the extent of the object being measured is much smaller than the extent of the reference, then the object cannot be found both before and after.
+Quantities over which we have an extent, like space and time, are another matter.
+When measuring the position of an object, this can extend before, on and after the mark.
+If the extent of the object being measured is much smaller than the extent of the reference, then we can assume the object to be wholly found either before, on or after.
 In those regimes we can treat the references as strict.
@@
 This is in line with the perspective of considering a physical theory, including quantum theory, as dealing with the output of measurements.
-For example, Wheeler commented: ``It is wrong, moreover, to regard this or that physical quantity as sitting out there with this or that numerical value''.[TODO: Add ref] What we are doing is going to a more fundamental level to understand how this works in the continuous case, and how can it be modeled in a rigorous say.
+For example, Wheeler commented\cite{Wheeler}: ``It is wrong, moreover, to regard this or that physical quantity as sitting out there with this or that numerical value''.
+What we are doing is going to a more fundamental level to understand how this works in the continuous case, and how can it be modeled in a rigorous say.
 
@@
 For (ii), we saw that references are objects that have, in general, an extent in space.
-A reference can be considered strict if the extent of the object measured is much smaller than the extent of the reference [TODO: Fix however gets fixed above].
-For example, the top level of a liquid in a jar is thinner than the marks used to measure the content.
-Ultimately, however, the finest constituents will also be the finest objects that we can use as finest references.
+A reference can be considered strict if the extent of the object measured is much smaller than the extent of the reference.
+Ultimately, however, the finest constituents will also be the finest objects that we can use as references.
 Therefore, if we assume that we can make finer and finer references, at some point we will find the extent of the references comparable to the extent of the objects we are measuring.
-This is also true for fundamental particles, where the extent of the system corresponds to the extent of the wave function: if the wavefunctions overlap, one particle is not clearly before/on/after the other.
+This is also true for fundamental particles, where the extent of the system corresponds to the extent of the wave function: if the wavefunctions overlap, one particle is not clearly wholly before/on/after the other.
 Similar arguments apply for excitations of quantum fields.
@@
 
+\section*{References}
+
+\bibliographystyle{unsrt}
+\bibliography{bibliography}
+
 \end{document}

### 2020-02-26 JustTheMath/probabilityTheory/probabilityTheory.tex
Adding JTM for probability

\documentclass{article}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
% Remove line spaces between items of enumerate and itemize
\usepackage{enumitem}
\setlist{noitemsep}

\newtheorem{assump}{Assumption}
\renewcommand*{\theassump}{\Roman{assump}}

\newtheorem{axiom}[equation]{Axiom}
\newtheorem{defn}[equation]{Definition}
\newtheorem{prop}[equation]{Proposition}
\newtheorem{coro}[equation]{Corollary}
\newtheorem{thrm}[equation]{Theorem}

%\theoremstyle{definition}

\newenvironment{remark}{\emph{Remark}.}{}
\newenvironment{rationale}{\emph{Rationale}.}{\qed}
\newenvironment{justification}{\emph{Justification}.}{\qed}
\renewenvironment{proof}{\emph{Proof}.}{\qed}

\title{Just the math: probability theory}

\date{\vspace{-5ex}}
\begin{document}

\maketitle

\begin{abstract}
Short summary of probability theory.
\end{abstract}

\section{Measure theory}

\begin{defn}
	Given a set $E$, a $\sigma$-algebra $\mathcal{F} \subseteq 2^E$ is a collection of subsets of $E$ such that it:
	\begin{itemize}
		\item contains the entire space (i.e. $E \in \mathcal{F}$)
		\item is closed under complement (i.e. if $A \in \mathcal{F}$ then $A^C \in \mathcal{F}$)
		\item is closed under countable union (i.e. if $\{A_i\}_{i=1}^\infty \subseteq \mathcal{F}$ then $\bigcup_{i=1}^\infty A_i \in \mathcal{F}$)
	\end{itemize}
\end{defn}

\begin{coro}
	By DeMorgan laws, a $\sigma$-algebra is also closed under countable intersections.
\end{coro}

\begin{defn}
	A \textbf{measurable space} is a tuple $(E, \mathcal{F})$ where $E$ is a set and $\mathcal{F}$ is a $\sigma$-algebra over $E$.
\end{defn}

\begin{defn}
	A \textbf{measure} over a measurable space $(E, \mathcal{F})$ is a function $\mu : \mathcal{F} \to [0, +\infty]$ such that:
	\begin{itemize}
	\item is \textbf{countably additive} (also called \textbf{$\sigma$-additive}) (i.e. $\mu(\bigcup_{i=1}^\infty A_i) = \sum_{i=1}^\infty \mu( A_i)$ where $\{A_i\}_{i=1}^\infty \subseteq \mathcal{F}$ is a countable collection of disjoint sets)
	\item the measure of the empty set is zero (i.e. $\mu(\emptyset) = 0$)
\end{itemize}
\end{defn}

\begin{defn}
	Given two measurable spaces $(A, \mathcal{F})$ and $(B, \mathcal{G})$, a \textbf{measurable function} is a map $X : A \to B$ such that the reverse image of an element of the $\sigma$-algebra of $B$ is in the $\sigma$-algebra of $A$ (i.e. $X^{-1}(U) \in \mathcal{F}$ where $U \in \mathcal{G}$).
\end{defn}

\section{Probability space}

\begin{defn}
	A \textbf{probability space} is a triple $(\Omega, \mathcal{F}, P)$ where:
	\begin{itemize}
		\item the set $\Omega$ is the \textbf{sample space}, or \textbf{possibility space}, that contains all the possible outcomes
		\item an \textbf{event} $A \subseteq \Omega$ is a set of possible outcomes and the $\sigma$-algebra $\mathcal{F}$ is the collection of all possible events
		\item the measure $P : \Omega \to [0,1]$ is the \textbf{probability measure} that returns one for the entire sample space (i.e. $P(\Omega) = 1$)
	\end{itemize}
\end{defn}

\section{Random variables}

\begin{defn}
	Given a measurable space $(E, \mathcal{G})$, a \textbf{random variable} $X : \Omega \to E$ is a measurable function. Typically, $E = \mathbb{R}$.
\end{defn}

The following definitions apply only over random variables over the reals.

\begin{defn}
	Given a random variable $X$, the \textbf{Cumulative Distribution Function} or \textbf{CDF} $F_X : \mathbb{R} \to [0,1]$ is given by $F_X(x) = P(X \leq x)$ or more precisely $F_X(x) = P(X^{-1}( (-\infty, x] ))$.
\end{defn}

\begin{coro}
	For all random variables:
	\begin{itemize}
		\item $F_X(-\infty) = 0$ and $F_X(+\infty) = 1$
		\item $F_X$ is monotonic
		\item $P(a < X \leq b) = F_X(b) - F_X(a)$ 
	\end{itemize} 
\end{coro}

\begin{defn}
	A random variable is \textbf{discrete} if the CDF is piecewise constant, is \textbf{continuous} if the CDF is continuous, is \textbf{mixed} if it is neither.
\end{defn}

\begin{defn}
	For a continuous random variable, the \textbf{probability density function} or \textbf{PDF} $f_X : \mathbb{R} \to \mathbb{R}$ is the Radon–Nikodym derivative of the CDF (i.e. $f_X = \frac{d F_x}{dx}$).
\end{defn}

\begin{remark}
	The Radon–Nikodym derivative is a purely measure theoretic way to define the derivative.
\end{remark}

\begin{coro}
	For all continuous random variables:
	\begin{itemize}
		\item $P[a \leq X \leq b] = \int_a^b f_X dx$
		\item $\int_{-\infty}^{\infty} f_X dx = 1$
	\end{itemize} 
\end{coro}

\begin{defn}
	Given a non-negative random variable $X$, the expectation value is defined to be $E[X] = \int_\Omega X dP(\omega)$. In general, $X = X^+ - X^-$ so we can write $E[X] = E[X^+] - E[X^-]$ if both 
	$$
	E[X]=
	\begin{cases}
	E[X^+] - E[X^-]  &   E[X^+] < \infty; E[X^-] < \infty \\
	\infty &   E[X^+] = \infty; E[X^-] < \infty \\
	-\infty &   E[X^+] < \infty; E[X^-] = \infty \\
	undefined &   E[X^+] = \infty; E[X^-] = \infty
	\end{cases}
	$$
\end{defn}

\begin{remark}
	The integral used is the Lebesgue integral.
\end{remark}

%\bibliographystyle{plain}
%\bibliography{bibliography}

\end{document}

### 2020-02-26 Proceedings/2020-QPL/2020QPL-ExtendedAbstract.tex
First draft at QPL abstract

\documentclass{article}
\usepackage{authblk}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}

% Remove line spaces between items of enumerate and itemize
\usepackage{enumitem}
\setlist{noitemsep}

% Adds double bracket symbols
\usepackage{stmaryrd}

% Latex symbol guide at http://mirrors.ibiblio.org/CTAN/info/symbols/comprehensive/symbols-letter.pdf

% LOGIC symbols
% -------------

% Allows to create negation symbols
\usepackage{MnSymbol}

% General math symbols
\DeclareMathOperator{\Id}{Id} % Identity function

% Boolean symbols and algebra
\def\Bool{\mathbb{B}}
\def\TRUE{\textsc{true}}
\def\FALSE{\textsc{false}}
\def\AND{\wedge}
\def\bigAND{\bigwedge}
\def\OR{\vee}
\def\bigOR{\bigvee}
\def\NOT{\neg}

% Logical context and related symbols
\def\logCtx{\mathcal{S}}
\def\vstmtSet{\mathcal{S}_\textsf{v}}
\def\dstmtSet{\mathcal{S}_\textsf{d}}
\newcommand{\pAss}[1][\mathcal{S}] {\mathcal{A}_{#1}}
\DeclareMathOperator{\truth}{truth}

% Experimental test symbols
\newcommand{\exptSet}{\mathcal{E}}
\newcommand{\expt}[1][e] {\mathsf{#1}}
\DeclareMathOperator{\result}{result}
\def\SUCCESS{\textsc{success}}
\def\FAILURE{\textsc{failure}}
\def\UNDEF{\textsc{undefined}}

% Statements
\def\tautology{\top} % Tautology
\def\contradiction{\bot} % Contradiction
\newcommand{\stmt}[1][s] {\mathsf{#1}} % Statement
\newcommand{\tstmt}[1][s] {\bar{\mathsf{#1}}} % Theoretical statement

% Relationships between statements
\def\comp{\doublefrown} % Compatibility
\def\ncomp{\ndoublefrown} 
\def\narrower{\preccurlyeq} % Narrowness
\def\nnarrower{\npreccurlyeq}
\def\snarrower{\prec}
\def\nsnarrower{\nprec}
\def\broader{\succcurlyeq} % Broadness
\def\nbroader{\nsucccurlyeq}
\def\sbroader{\succ}
\def\nsbroader{\nsucc}
\def\indep{\upmodels} % Independent
\def\nindep{\nupmodels}

% Experimental domains and related symbols
\newcommand{\edomain}[1][D] {\mathcal{#1}} % Experimental domain
\newcommand{\tdomain}[1][D] {\bar{\mathcal{#1}}} % Theoretical domain
\newcommand{\basis}[1][B] {\mathcal{#1}} % Basis
\newcommand{\resPoss}[1][x] {\mathring{#1}} % Residual possibility
\newcommand{\estPoss}[1][x] {\dot{#1}} % Established possibility

% Formatting for experimental relationships
\newcommand{\erel}[1][r] {#1}

% Formatting for sentence statements
\newcommand{\statement}[1] {\emph{``#1"}}

% Formatting for reference
\newcommand{\refStmt}[1][r]{\textbf{#1}}

\DeclareMathOperator{\ver}{ver}
\DeclareMathOperator{\fal}{fal}
\DeclareMathOperator{\und}{und}

\DeclareMathOperator{\interior}{int}
\DeclareMathOperator{\exterior}{ext}

\usepackage{xcolor} % Required for specifying colors by name
\definecolor{sectionNumbers}{RGB}{44, 103, 0}

\renewcommand\thesubsection{\thesection.\Alph{subsection}}
\renewcommand{\theequation}{\thesection.\arabic{equation}}

\newtheorem{assump}{Assumption}
\renewcommand*{\theassump}{\Roman{assump}}

\newtheorem{axiom}[equation]{Axiom}
\newtheorem{defn}[equation]{Definition}
\newtheorem{prop}[equation]{Proposition}
\newtheorem{coro}[equation]{Corollary}
\newtheorem{thrm}[equation]{Theorem}

%\theoremstyle{definition}

\newenvironment{remark}{\emph{Remark}.}{}
\newenvironment{rationale}{\emph{Rationale}.}{\qed}
\newenvironment{justification}{\emph{Justification}.}{\qed}
\renewenvironment{proof}{\emph{Proof}.}{\qed}

\title{A model for the experimental verification of the continuum  OR \\ The continuum from experimental verification OR \\ A model of the continuum from experimental verification OR \\ Experimental verification and the continuum}
\author[1]{Gabriele Carcassi}
\author[1]{Christine A. Aidala}
\affil[1]{
	Physics Department\\
	University of Michigan
	}

\date{\today}
\begin{document}

\maketitle

\begin{abstract}
We present a model for the continuum that starts from requirements of experimental verification. That is, instead of starting from abstract mathematical structures or constructions, we posit that we can use a physical object as a reference such that we are able to verify whether another object lies before or after. We then lay out a series of necessary and sufficient conditions such that the experimentally distinguishable cases defined by a set of references can be labeled with a real number. We find that most of the conditions are needed to recover a more general notion of ordering, while the additional conditions to recover the reals (or the integers) are minimal. We argue that, at Planck scale, the conditions for ordering are not tenable. How and why this is may give new insights to people interested in space-time structure.
\end{abstract}

\section{Introduction}

In our ongoing project, Assumptions of Physics, we aim to find a set of minimal assumptions from which the different physical theories can be derived. By doing this as formally as possible, we are forced to specify all our implicit starting points, which provides greater insight and new ideas for solving open problems. While other approaches have a similar goal, see for example Refs.~\cite{PhysRevA.84.012311,QLogicReview,Hardy:2001jk,ludwig_hein_2013}, there are a number of key differences in our approach.

The first difference is that we want the different theories (e.g. classical Hamiltoniam mechanics, quantum particle mechanics, thermodynamics, ...) to be derived from starting points that are immediately intuitive to the non-expert. The assumptions we work with describe, for example, whether the state of the whole system is equivalent to the state of its parts (i.e. reducibility), or whether knowledge of the future state is equivalent to knowledge of the past state and vice-versa (i.e. determinism and reversibility). Physically, the idea is that we have a system to study and, under a few simplifying assumptions, we find that the system is adequately described by one of the theory. The set of assumptions for each theory, then, work both as a starting point to rederive the theory and as a clear demarcation of its realm of applicability. For example, Hamiltonian mechanics is derived on the assumption of determinism and reversibility, therefore we know it will not work if our system is irreversible.

Another difference is that our approach aims to start with primitives that are necessary to do physics, and not with requiring abstract mathematical structures that, at the very end, will give us something that is interpretable as physics. For example, one of our basic building block is the notion of a verifiable statement: an assertion for which an experimental test is available that would confirm, in finite time, that the assertion is true. As one can imagine, without such a notion, we would have no experimental verification, which means no possibility of doing science. In general, the goal is to start with a tight connection between the physical concepts and their respective formal representations, so that it is always clear what the mathematical structure represents physically and no mathematical objects are unphysical. Ideally, this would tell us, for example, what a topology or a differentiable structure represents physically in current theories, and when those are valid approximations of our physical concepts. In this particular work, the goal is to provide that understanding for real valued quantities.

The last difference we want to highlight is that we typically start at a lower level. For example, we do not assume probability spaces, which are already constructions upon sigma-algebras and measures, and sigma-algebras are typically constructed from the topology of the space. We instead aim to recover those as well. This is particularly important if one is interested in the structure of space-time, since its topology and sigma-algebra are the most foundational aspects.

What we want to present here is a construction of continuous quantities from the basic notion of verifiable statements, assertion that can be verified experimentally in finite time. We believe this is of interest to those working on the fine structure of space-time as it gives us necessary and sufficient operational requirements such that a set of experimentally distinguishable cases can be labeled by real numbers. The basic building block to construct quantities is the notion of a reference: a physical object that allows to distinguish between a before and an after. These generalize the notion of ticks of a ruler, pulses within a timing system, reference weights for a balance scale. The notion of a quantity, then, is not something that exists a priori, but something that is constructed through a set of prepared references, with known logical relationships among them. The goal, then, is to understand what are the necessary and sufficient conditions that a set of references must satisfy so that they distinguish a set of cases that can be labeled by real numbers.

Note that the set of real numbers can be characterized in different ways since they are usually associated with multiple structures. It is important to stress that, in our case, we are interested in the most basic characterization, which is in terms of ordering. Any totally ordered set that is dense, complete and contains a countable dense subset is order isomorphic to the real numbers. One can, through a non-linear monotonic transformation, change the group/field structure while preserving the ordering, but it is not possible to do the opposite. In this sense, the ordering structure is more fundamental than the group/field structure. Also note that the ordering is exactly enough to define the topology and nothing more (i.e. no geometric information).

The first main result is that much of the requirements are needed simply to impose an ordering. Comparatively, the requirements to turn that ordering into the one of the continuum are not as involved. The second main result is that the requirements for ordering are such that they are most likely not realized at Planck scale, leading to a breakdown not just of real valued quantities, but essentially of any scalar numeric quantity and, consequently, the mathematical techniques that rely on them. While we do not have yet a clear idea how to proceed in those cases, the work forced us to develop techniques to construct the ordering from more fundamental notions. The essential difficulty was to find concepts that are still physically intuitive that do not implicitly lead to circular arguments (e.g. any pictorial representation is inherently ``ordered" and should be avoided). We believe these may be useful to other as they may give new ideas and insights.

On the technical level, the full details are available at \cite{Carc3} which effectively function as a large appendix. Here we provide a summary of the mathematical definitions and the results, skipping much of the details, in particular the proofs and the parts that are not strictly necessary for our purposes.

\section{Logical contexts and experimental domain}

Our starting point is that a physical theory consists, at the very least, of a set of statements with well defined logical relationships defined by the semantic of the theory. We leave open the question of how those are constructed or articulated, but a set of logically consistent statement must be given. Some, but not all, the statements are experimentally verifiable, that is we have at our disposal a test that, if the statement is true, the test terminates successfully in finite time. We again leave open the question of how that is done and what constitutes enough experimental evidence, but a set of experimental tests must be given. For example, \statement{the mass of the particle is less than 1 keV} can be considered verifiable while \statement{the mass of the particle is exactly 1 keV} cannot, since we cannot perform infinite precision measuremnets.

The semantic of a specific statement (e.g. \statement{the mass of the electron is 0.5 $\pm$ 0.05 MeV} ) may in general depend on the definitions of the different terms and the characterization of the experimental procedures. As such, our most fundamental object is a set $\logCtx$ called \textbf{logical context} and each element $\stmt \in \logCtx$ is a \textbf{statement}. Each logical context is equipped with a set $\pAss \subseteq \Bool^\logCtx$, where $\Bool = \{ \TRUE, \FALSE \}$. Each $a \in \pAss$ represents a \textbf{possible assignment}, a possible way to set the truth of each value that is consistent with the semantic of the statements. For example, there will be no possible assignment where \statement{the particle is an electron} and \statement{the particle is a photon} are true at the same time.

This basic structure allows us to define:
\begin{itemize}
	\item \textbf{tautologies} and \textbf{contradictions}, noted $\tautology$ and $\contradiction$, as statements that are respectively always true or always false in all possible assignments (i.e. for all $a \in \pAss$, $a(\tautology) = \TRUE$ and $a(\contradiction) = \FALSE$)
	\item a notion of \textbf{statement equivalence}, noted $\stmt_1 \equiv \stmt_2$, if two statements have the same value in all possible assignments (i.e. for all $a \in \pAss$, $a(\stmt_1) = a(\stmt_2)$)
	\item a notion of \textbf{narrowness}, noted $\stmt_1 \narrower \stmt_2$, if the second statement is true whenever the first one is (i.e. for all $a \in \pAss$ where $a(\stmt_1) = \TRUE$ we have $a(\stmt_2) = \TRUE$); the opposite notion is called \textbf{broadness} 
	\item a notion of \textbf{compatibility}, noted $\stmt_1 \comp \stmt_2$, if the statements can be true at the same time (i.e. there exists an $a \in \pAss$ such that where $a(\stmt_1) = a(\stmt_2) = \TRUE$)
\end{itemize}
We can also define a notion of \textbf{independence} on a set of statement $S$, noted $\stmt_1 \indep \stmt_2$ for a binary set, if fixing the truth value of any subset does not constrain the truth value of the others, though it will not be used here.

Furthermore, the set of equivalence classes $\logCtx_{/\equiv}$ will form a complete Boolean algebra with the standard negation/conjunction/disjunction operations and properties, which also means $(\logCtx_{/\equiv}, \narrower)$ is an ordered set. This structure captures the idea that a physical theory is a set of logical statements with well defined logical relationships. From now on, unless otherwise stated, statement will refer to its equivalence class.

To keep track of experimental verifiability, every logical content $\logCtx$ comes equipped with a subset $\vstmtSet \subseteq \logCtx$, and each element $\stmt \in \vstmtSet$ will be called a \textbf{verifiable statement}. The idea is that, for each verifiable statement, we have at our disposal an experimental test that, if the statement is true, the test terminates in finite time. The requirement of experimental verifiability over finite time imposes the following conditions:
\begin{itemize}
	\item all tautology and contradictions are verifiable, since the context assumes them to be always true or false
	\item a statement equivalent to a verifiable statement is verifiable, since we can use the test of the second to test the first
	\item the finite conjunction and the countable disjunction of verifiable statements is verifiable, since we can construct general algorithms for both cases that simply execute the original tests in succession appropriately
\end{itemize}
Note that $(\vstmtSet, \narrower)$ is also a lattice, and is embedded in $(\logCtx, \narrower)$. This double structures allows us to keep track of what is logically consistent and what is experimentally verifiable within a theory.

Finite time experimental verification also imposes a limit on how many statements one can verify. Given an arbitrarily long amount of time, a countable set of verifiable statement represents the maximum. Given a set $S \subseteq \vstmtSet$ of verifiable statements, we call a \textbf{basis} $\basis \subset S$ a subset from which all of $S$ can be generated through finite conjunction and countable disjunction. We define an \textbf{experimental domain} $\edomain_X \subseteq \vstmtSet$ a set of verifiable statements that includes the tautology and the contradiction, that is closed under finite conjunction and countable disjunction and that can be generated by a countable basis. An experimental domain represents the most that can be learned experimentally about a particular scientific problem. Note that an experimental domain is a Heyting algebra.

An experimental domain will contain the negation of a verifiable statement only if its relative test always terminates. Even if the test does not always terminate in the negative case, though, it still conceptually creates a possible case within the context. Given an experimental domain $\edomain_X$ we construct its \textbf{theoretical domain} $\tdomain_X$ by closing under negation, as well as the original finite conjunction and countable disjunction. This represents the set of statements for which an experimental test is well defined, regardless of whether it will terminate or not. By DeMorgan rules, the theoretical domain is closed under countable conjunction as well, so it is a $\sigma$-complete Boolean algebra.

Within the theoretical domain, we define the  \textbf{possibilities of the domain} $X = \{ x \in \edomain_X \, | \, \forall \tstmt \in \tdomain_X $ either $x \narrower \tstmt$ or $x \ncomp \tstmt\}$ as those statements that, if true, will define the truth value for all other statements in the theoretical domain. These correspond to the cases that are experimentally distinguishable through the verifiable statements in the experimental domain. Conceptually, each possibility corresponds to a unique possible assignment for the domain. Mathematically, each possibility is a minterm of the basis (i.e. a conjunction where each element of the basis appears exactly once, either negated or not) that is not a contradiction.

Since every statement of the experimental domain can be expressed as a disjunction of minterms of the basis, for each verifiable statement $\stmt \in \edomain_X$ we can find a set $U \subseteq X$ such that $\stmt = \bigOR_{x \in U} x$. Similarly, for each theoretical statement $\tstmt \in \tdomain_X$ we can find a set $A \subseteq X$ such that $\tstmt = \bigOR_{x \in A} x$. Logical operations on the statements become set operations on the sets, which means that an experimental domain induces a natural topology on its possibilities, while a theoretical domain induces a natural $\sigma$-algebra, which is the Borel algebra of the natural topology.

This basic structure gives a well defined map between the physical objects and their mathematical counterparts. For example, given a theoretical statements $\tstmt$ we can define its verifiable part $\ver(\tstmt)$ as the broadest statement that is narrower than $\tstmt$. This corresponds to the cases in which the statement is true and the test terminates. If a statement is verifiable, then $\ver(\tstmt) \equiv \tstmt$. Topologically, this corresponds to the interior of a Borel set and, indeed, an open set corresponds to its interior. Conversely, we can define a falsifiable part $\fal(\tstmt)$ which, topologically, corresponds to the exterior of a Borel set. A statement that has nor interior or exterior, like the set of all rationals over a real line, corresponds to a statement for which the test never terminates: an undecidable statement. One can proceed this way and re-understand all definitions and proofs.

The general result is that any set $X$ of experimentally distinguishable cases has to come with a $\mathsf{T}_0$ second countable topology that represents the verifiable statements and a Borel algebra that represents the theoretical statements. These structures are the foundation for most of the other mathematical structures used in science (e.g. differential topology, measure theory, probability theory, ...) and this framework provides a strong justification for them and a direct physical meaning. It also gives us an understanding of exactly why those requirements are needed. A topological space that is not second countable, for example, would correspond to a set of possibility that cannot be, even in the limit, experimentally distinguished. A set with cardinality greater then that of the continuum, for example, does not allows a second countable topology and cannot correspond to a set of experimentally distinguishable cases: each element, in fact, must be identified with the output of a countable binary tests, which roughly corresponds to a real number in base two. A non-Borel set is also non-physical as it does not correspond to a statement for which we can construct a test, regardless of whether it terminates. In our view, much confusion is there in physics because we do not have such precise and methodical mapping between physical ideas and mathematical ones.

To sum up, we start with an experimental domain $\edomain_X$ which can be thought as a topology/locale/Heyting algebra with a countable basis. We close under negation to obtain a theoretical domain $\tdomain_X$, which corresponds to the related Borel algebra, and there find the points of our space $X$, which represent the physically distinguishable cases. We again stress that the points are generated from the verifiable statements. This framework allows us to ask the question: when is a set of physically distinguishable cases identifiable with the real numbers? Under what conditions the points become isomorphic to those of a real line?

\section{References and experimental ordering}

In general, we can concentrate on a single property of an object and measure that independently. It is fairly typical to define a quantity as an ordered property, where different values can be compared to see which one is greater. We want to capture these notions in our framework.

If we concentrate on a single property of an object, we can select the verifiable statement that only pertain to that property. These will form an experimental domain $\edomain_X$ whose possibilities $X$ will correspond to all possible property values. Recall that, given a totally ordered set $(\mathcal{Q}, \leq)$, the \textbf{order topology} is the one generated by sets of the form $$(a, \infty) = \{q \in \mathcal{Q} \, | \, a < q\} \;,\; (-\infty, b) = \{q \in \mathcal{Q} \, | \, q < b\}.$$
The standard topologies for the integers and the reals are simply the order topologies with their respective ordering. We will say that an experimental domain is \textbf{naturally ordered} when the possibilities are totally ordered and the natural topology is the corresponding order topology. Our goal, then, is to understand when the logical relationships between verifiable statements for a particular property are such that the possibilities are naturally ordered like the real numbers.

First note that the basis of the order topology corresponds to verifiable statements of the form \statement{the property value is after $a$} and \statement{the property value is before $b$}. This parallels what happens in practice: in general we have a set of references (e.g. marks on a ruler or ticks of a clock) which ideally define known values of the quantity, and we are able to test whether the object is before or after that reference value. We say ideally because, in general, a reference does not span a single value. For example, a mark on a ruler will also have an extent in space. Also note that, in general, the object being measured can also have an extent. For example, an event may start before the tick of a clock and finish after. Before and after, then, are not necessarily mutually exclusive. Our most general definitions, then, need to account for these cases as well.

Let us define a \textbf{reference} $\refStmt = ( \stmt[b], \stmt[o], \stmt[a] )$ as a tuple of three statements (the before/on/after statements) such that:
\begin{enumerate}
	\item we can verify whether the object is before or after the reference: $\stmt[b]$ and $\stmt[a]$ are verifiable statements
	\item the reference itself must occupy some values : $\stmt[o] \nequiv \contradiction$
	\item if the object is not before or after, it is on the reference: $\NOT \stmt[b] \AND \NOT \stmt[a] \narrower \stmt[o]$
	\item if the object is before and after, it is also on the reference: $\stmt[b] \AND \stmt[a] \narrower \stmt[o]$
\end{enumerate}
A set of references generates an experimental domain in the following way. Let $R = \{( \stmt[b]_i, \stmt[o]_i, \stmt[a]_i )\}_{i \in I}$ be a set of references. Let $\basis_b = \{\stmt[b]_i\}_{i \in I}$ be the set of all before statements and $\basis_a = \{\stmt[a]_i\}_{i \in I}$ be the set of all after statements. The experimental domain $\edomain_X$ generated by $R$ is the one generated by all before and after statements $\basis_b \cup \basis_a$. The \textbf{before domain} $\edomain_b$ is the domain generated only by the before statements $\basis_b$ and the \textbf{after domain} $\edomain_a$ is the domain generated only by the after statements $\basis_a$. The possibilities $X$ of the domain generated by $R$ consists of all the cases that are experimentally distinguishable given the references.

We can see that if $(X, \leq)$ is a totally ordered set of mutually incompatible statements (i.e. no two can be true at the same time) whose union is a tautology (i.e. one of them must be true), we can create the collection of triples $R = \{ \bigOR\limits_{x_b < x} x_b , x , \bigOR\limits_{x_a > x} x_a \}_{x \in X}$. If the before and after statements are verifiable, then the triples are references and the natural topology of the experimental domain they generate is exactly the order topology. Therefore, every naturally ordered experimental domains can be generated by a set of references. What we are looking is for the opposite conditions: what are the minimum requirements on a set of references such that they generate a naturally ordered experimental domain?

\subsection*{Requirements for ordering}

One of the main guiding insights is that, in a total order, $(a, +\infty) ^C = (-\infty, a]$. If $a \geq b$ then we have $(-\infty, b) \subseteq (-\infty, a]$ and if $a < b$ then $(-\infty, a) \subset (-\infty, b]$. This means that if $X$ is totally ordered, then all the sets of them form $(-\infty, x)$ and the complements of $(x, \infty)$ will be ordered by set inclusion. In terms of statement, this means that all the before statements $\basis_b$ and the negation of the after statements $\NOT(\basis_a)$ are ordered by narrowness. This, turns out, is a necessary condition.

We say a reference is \textbf{strict} if the before/on/after cases are mutually exclusive. That is, $\refStmt = (\stmt[b], \stmt[o], \stmt[a])$ is such that $\stmt[b] \ncomp \stmt[a]$ and $\stmt[o] \equiv \NOT \stmt[b] \AND \NOT \stmt[a]$. For a strict reference we then have $\stmt[b] \narrower \NOT \stmt[a]$. This is automatically achieved for discrete quantities, since object do not have an extent over them. For those that have an extent, for example when measuring the position of an object, this assumes that the extent of the object to be measured is very small compared to the extent of the reference so that the object can always be considered wholly before/on/after the reference.

We say two references $\refStmt_1 = (\stmt[b]_1, \stmt[o]_1, \stmt[a]_1)$ and $\refStmt_2 = (\stmt[b]_2, \stmt[o]_2, \stmt[a]_2)$ are \textbf{aligned} if for any $\stmt_1 \in \{ \stmt[b]_1, \NOT\stmt[a]_1\}$ and $\stmt_2 \in \{ \stmt[b]_2, \NOT\stmt[a]_2\}$ we have 
$\stmt_1 \narrower \stmt_2$ or $\stmt_2 \narrower \stmt_1$. A set $R$ of references is aligned if any pair of references is aligned. If two references are aligned and do not overlap (i.e. $\stmt[o]_1 \ncomp \stmt[o]_2$) then one reference must be before the other. Formally, one reference is before the other, noted $\refStmt_1 \less \refStmt_2$, if $\stmt[b]_1 \OR \stmt[o]_1 \ncomp \stmt[o]_2 \OR \stmt[a]_2$. If they do not overlap, they at least are such that we can imagine they have boundaries that come one after the other. Moreover, alignment implicitly requires that we can always tell references apart such that their relationship are always expected to be the same. If they jiggle around there can be no fixed, well-defined, alignment.

Another insight is that $(-\infty, b) \subset (-\infty, b]$ and no other set of the form $(-\infty, x)$ or $(-\infty, x]$ can be placed in between. In terms of statement, this means that the totally ordered set $( \basis_b \cup \NOT(\basis_a), \narrower)$ is formed by pairs of consecutive elements, one for each possible value.

We say a set of references $R$ is \textbf{refinable} if given two strict references $\refStmt_1 = ( \stmt[b]_1, \stmt[o]_1, \stmt[a]_1)$ and $\refStmt_2 = ( \stmt[b]_2, \stmt[o]_2, \stmt[a]_2)$ we can always:
\begin{itemize}
	\item find an intermediate one if they are not consecutive; that is, if $\refStmt[r]_1 < \refStmt[r]_2$ but $\refStmt[r]_2$ is not the immediate successor of $\refStmt[r]_1$ (i.e. $\NOT \stmt[a]_1 \snarrower \stmt[b]_2$), then we can find a strict reference $\refStmt_3$ aligned with $\edomain$ such that $\refStmt[r]_1 < \refStmt[r]_3 < \refStmt[r]_2$.
	\item refine overlapping references if one is finer than the other; that is, if $\stmt[o]_2 \snarrower \stmt[o]_1$, we can find a strict reference $\refStmt_3$ aligned with $\edomain$ such that $\stmt[o]_3 \narrower \stmt[o]_1$ and either $\stmt[b]_3 \equiv \stmt[b]_1$ and $\refStmt_3 < \refStmt_2$ or $\stmt[a]_3 \equiv \stmt[a]_1$ and $\refStmt_2 < \refStmt_3$.
\end{itemize}
This conditions means that we are able to cover the whole space with references fine enough to resolve all possible overlaps. This is what allows us to pick out individual values as we can ultimately find a before statement that is broadest narrower statement of each negated after statement.

With the given definitions, we are able to state the following three theorems, the proofs of which are available at \cite{Carc3}.
\begin{thrm}[Reference ordering theorem]
	An experimental domain is naturally ordered if and only if it can be generated by a set of refinable aligned strict references.
\end{thrm}
\begin{thrm}[Discrete ordering theorem]
	An experimental domain is naturally ordered with a discrete quantity if and only if it is generated by a set of refinable aligned strict references with a sparse order (i.e. only finitely many references between two)
\end{thrm}
\begin{thrm}[Continuous ordering theorem]
	An experimental domain is naturally ordered by a continuous quantity if and only if it is generated by a set of refinable aligned strict references with a dense order (i.e. infinitely many references between two)
\end{thrm}

Note that the only extra condition to recover the continuous and the discrete orderings is the cardinality of references found between two. This may be unexpected for the continuous case, since the real numbers not only require the ordering be dense, but also complete and to have a countable dense subset. What happens is that the experimental domain automatically provides the other two conditions automatically. The closure of the domain under countable disjunction requires the ordering to be complete. The countable basis provides the countable dense subset. Therefore the dense ordering is the only extra condition.

The other interesting aspect is that in the discrete order all the before and after statements are not only verifiable, but also falsifiable. In fact, this is the only ordering with that characteristic. On the other hand, in the continuous order none of the before and after statements are falsifiable. In fact, this is the only ordering with that characteristic. This stresses that, experimentally, discrete and continuous quantities are qualitatively different.

\subsection*{At finer scale}

Now that we have understood what are the requirements to experimentally define a continuous quantity, we can ask the following question: as we increase spatial and temporal resolution and we approach Plank scale, can we still satisfy those conditions? Or can they only be satisfied as an approximation at large enough scale?

At this point, we do not aim lay out a convincing argument for or against. Rather, we want to show what are the main points, both conceptually and technically, that one may raise so that one can understand how to construct his own argument. We believe this to be more interesting because, once one decides how and why those assumptions fail, he is already on the path to understand what alternative he wants to explore. We are going to reason in terms of space, but similar arguments can be made in terms of time.

One of the requirements is that references are strict, that is the before/on/after cases are mutually exclusive. When measuring position in space, where finite objects have a finite extent, those cases are not in general mutually exclusive because of overlaps. We can ignore the problem by assuming that the object whose position we want to measure is much smaller than the reference, in which case we can assume that the object is wholly before/on/after the reference. But in order to derive continuous quantities, we also assume that we can take finer and finer references. If the objects we use as references of position are the same objects of which we want to measure the position we logically come at an impasse. This is going to be a problem for particles (i.e. fields excitation) as well since the overlap can be in terms of their wave-packet. In these cases, it is not clear one can make the claim that a particle will always be wholly before/on/after the other.

Another requirement is that references are aligned, that is we can establish clear before/after relationship between the edges of the references. One issue here is that we always need to be able to distinguish between two references such that fixed ordering relationship can be established. If the finest references are individual particles, and these are indistinguishable from each other, it is not clear how this is going to work. Note that we cannot use position itself to distinguish them because we run into circular arguments: we want to use the reference to distinguish position not the other way around. Another problem is that we cannot talk about a substructure for individual particles, therefore it is not clear whether the edges (i.e. where the particle ``starts") can be well defined if particles are used as references.

Another requirement is that references are refinable, that is we can always find finer references that resolve the overlap or be placed in gaps. While it is not necessary to have them all in the same setup as we can, in principle, repeat the experiment with different references in place to get different information, we still need to be able to construct them. While it is generally accept that there is a limit to how small a wave-packet can be made, meaning how small the extent of an object (or a reference) can be, there is no such limit between the distance between two objects. Therefore we can imagine creating situations in which two references are consecutive and yet no other reference can fit in between; or two references overlap yet no other reference can refine the overlap.

The punchline here is that if a single argument along these lines is accepted, the result is that we lose not just the real numbers, we lose ordering itself. What would happen is that, at short scales, ordering becomes fuzzy. Before and after, in both space and time, is not a sharp distinction. At least, we would be unable to define it in a way that is experimentally verifiable. With a loss of ordering, there is not a clear way to keep a sharp notion of distance and all other geometrical ideas. What we would have in its place would depend on how the structure breaks down and what new assumptions can be taken.

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

### 2020-02-29 Proceedings/2020-QPL/2020QPL-ExtendedAbstract.tex
Added bibliography

@@
 
-\title{A model for the experimental verification of the continuum  OR \\ The continuum from experimental verification OR \\ A model of the continuum from experimental verification OR \\ Experimental verification and the continuum}
+\title{A model of the continuum from experimental verifiability}
 \author[1]{Gabriele Carcassi}
@@
 \begin{abstract}
-We present a model for the continuum that starts from requirements of experimental verification.
+We present a model of the continuum that starts from requirements of experimental verifiability.
 That is, instead of starting from abstract mathematical structures or constructions, we posit that we can use a physical object as a reference such that we are able to verify whether another object lies before or after.
-We then lay out a series of necessary and sufficient conditions such that the experimentally distinguishable cases defined by a set of references can be labeled with a real number.
+We then lay out a series of necessary and sufficient conditions, in terms of logical relationships between experimentally verifiable statements, such that the experimentally distinguishable cases defined by a set of references can be labeled with a real number.
 We find that most of the conditions are needed to recover a more general notion of ordering, while the additional conditions to recover the reals (or the integers) are minimal.
-We argue that, at Planck scale, the conditions for ordering are not tenable.
-How and why this is may give new insights to people interested in space-time structure.
+We argue that, at Planck scale, the conditions for ordering are not tenable, which may give new insights to people interested in space-time structure at the finest scales.
 \end{abstract}
@@
 determinism and reversibility).
-Physically, the idea is that we have a system to study and, under a few simplifying assumptions, we find that the system is adequately described by one of the theory.
-The set of assumptions for each theory, then, work both as a starting point to rederive the theory and as a clear demarcation of its realm of applicability.
+Physically, the idea is that we have a system to study and, under a few simplifying assumptions, we find that the system is adequately described by one of the theories.
+The set of assumptions for each theory, then, works both as a starting point to rederive the theory and as a clear demarcation of its realm of applicability.
 For example, Hamiltonian mechanics is derived on the assumption of determinism and reversibility, therefore we know it will not work if our system is irreversible.
@@
 Another difference is that our approach aims to start with primitives that are necessary to do physics, and not with requiring abstract mathematical structures that, at the very end, will give us something that is interpretable as physics.
-For example, one of our basic building block is the notion of a verifiable statement: an assertion for which an experimental test is available that would confirm, in finite time, that the assertion is true.
+For example, one of our basic building blocks is the notion of a verifiable statement\cite{Carc2}: an assertion for which an experimental test is available that would confirm, in finite time, that the assertion is true.
 As one can imagine, without such a notion, we would have no experimental verification, which means no possibility of doing science.
@@
 
-What we want to present here is a construction of continuous quantities from the basic notion of verifiable statements, assertion that can be verified experimentally in finite time.
+What we want to present here is a construction of continuous quantities from the basic notion of verifiable statements, assertions that can be verified experimentally in finite time.
 We believe this is of interest to those working on the fine structure of space-time as it gives us necessary and sufficient operational requirements such that a set of experimentally distinguishable cases can be labeled by real numbers.
-The basic building block to construct quantities is the notion of a reference: a physical object that allows to distinguish between a before and an after.
+The basic building block to construct quantities is the notion of a reference: a physical object that allows one to distinguish between a before and an after.
 These generalize the notion of ticks of a ruler, pulses within a timing system, reference weights for a balance scale.
@@
 
-The first main result is that much of the requirements are needed simply to impose an ordering.
+The first main result is that most of the requirements are needed simply to impose an ordering.
 Comparatively, the requirements to turn that ordering into the one of the continuum are not as involved.
@@
 any pictorial representation is inherently ``ordered" and should be avoided).
-We believe these may be useful to other as they may give new ideas and insights.
+We believe these may be useful to others as they may give new ideas and insights.
 
-On the technical level, the full details are available at \cite{Carc3} which effectively function as a large appendix.
-Here we provide a summary of the mathematical definitions and the results, skipping much of the details, in particular the proofs and the parts that are not strictly necessary for our purposes.
+On the technical level, the full details are available at \cite{Carc3} which effectively functions as a large appendix.
+Here we provide a summary of the mathematical definitions and the results, skipping many of the details, in particular the proofs and the parts that are not strictly necessary for our purposes.
 
-\section{Logical contexts and experimental domain}
+\section{Logical contexts and experimental domains}
 
-Our starting point is that a physical theory consists, at the very least, of a set of statements with well defined logical relationships defined by the semantic of the theory.
-We leave open the question of how those are constructed or articulated, but a set of logically consistent statement must be given.
-Some, but not all, the statements are experimentally verifiable, that is we have at our disposal a test that, if the statement is true, the test terminates successfully in finite time.
-We again leave open the question of how that is done and what constitutes enough experimental evidence, but a set of experimental tests must be given.
+Our starting point is that a physical theory consists, at the very least, of a set of statements with well defined logical relationships defined by the semantics of the theory.
+We leave open the question of how those are constructed or articulated, but a set of logically consistent statements must be given.
+Some, but not all, the statements are experimentally verifiable, that is we have at our disposal a test that, if the statement is true, terminates successfully in finite time.
+We again leave open the question of how that is done and what constitutes enough experimental evidence, but a set of potential experimental procedures must be given.
 For example, \statement{the mass of the particle is less than 1 keV} can be considered verifiable while \statement{the mass of the particle is exactly 1 keV} cannot, since we cannot perform infinite precision measuremnets.
 
-The semantic of a specific statement (e.g.
-\statement{the mass of the electron is 0.5 $\pm$ 0.05 MeV} ) may in general depend on the definitions of the different terms and the characterization of the experimental procedures.
+The semantics of a specific statement (e.g.~\statement{the mass of the electron is 0.5 $\pm$ 0.05 MeV} ) may in general depend on the definitions of the different terms and the characterization of the experimental procedures.
 As such, our most fundamental object is a set $\logCtx$ called \textbf{logical context} and each element $\stmt \in \logCtx$ is a \textbf{statement}.
 Each logical context is equipped with a set $\pAss \subseteq \Bool^\logCtx$, where $\Bool = \{ \TRUE, \FALSE \}$.
-Each $a \in \pAss$ represents a \textbf{possible assignment}, a possible way to set the truth of each value that is consistent with the semantic of the statements.
+Each $a \in \pAss$ represents a \textbf{possible assignment}, a possible way to set the truth of each statement that is consistent with the semantics of the statements.
 For example, there will be no possible assignment where \statement{the particle is an electron} and \statement{the particle is a photon} are true at the same time.
@@
 	\item a notion of \textbf{compatibility}, noted $\stmt_1 \comp \stmt_2$, if the statements can be true at the same time (i.e.
-there exists an $a \in \pAss$ such that where $a(\stmt_1) = a(\stmt_2) = \TRUE$)
+there exists an $a \in \pAss$ such that $a(\stmt_1) = a(\stmt_2) = \TRUE$)
 \end{itemize}
-We can also define a notion of \textbf{independence} on a set of statement $S$, noted $\stmt_1 \indep \stmt_2$ for a binary set, if fixing the truth value of any subset does not constrain the truth value of the others, though it will not be used here.
+We can also define a notion of \textbf{independence} on a set of statements $S$ if fixing the truth value of any subset does not constrain the truth value of the others, though it will not be play a significant role here.
 
-Furthermore, the set of equivalence classes $\logCtx_{/\equiv}$ will form a complete Boolean algebra with the standard negation/conjunction/disjunction operations and properties, which also means $(\logCtx_{/\equiv}, \narrower)$ is an ordered set.
+Furthermore, the set of equivalence classes $\logCtx_{/\equiv}$ will form a complete Boolean algebra with the standard negation/conjunction/disjunction operations and properties, which also means $(\logCtx_{/\equiv}, \narrower)$ is a partially ordered set.
 This structure captures the idea that a physical theory is a set of logical statements with well defined logical relationships.
-From now on, unless otherwise stated, statement will refer to its equivalence class.
+From now on, unless otherwise stated, a statement will refer to its equivalence class.
 
-To keep track of experimental verifiability, every logical content $\logCtx$ comes equipped with a subset $\vstmtSet \subseteq \logCtx$, and each element $\stmt \in \vstmtSet$ will be called a \textbf{verifiable statement}.
-The idea is that, for each verifiable statement, we have at our disposal an experimental test that, if the statement is true, the test terminates in finite time.
+To keep track of experimental verifiability, every logical context $\logCtx$ comes equipped with a subset $\vstmtSet \subseteq \logCtx$, and each element $\stmt \in \vstmtSet$ will be called a \textbf{verifiable statement}.
+The idea is that, for each verifiable statement, we have at our disposal an experimental test that, if the statement is true, terminates in finite time.
 The requirement of experimental verifiability over finite time imposes the following conditions:
 \begin{itemize}
-	\item all tautology and contradictions are verifiable, since the context assumes them to be always true or false
+	\item all tautologies and contradictions are verifiable, since the context assumes them to be always true or false
 	\item a statement equivalent to a verifiable statement is verifiable, since we can use the test of the second to test the first
@@
 Note that $(\vstmtSet, \narrower)$ is also a lattice, and is embedded in $(\logCtx, \narrower)$.
-This double structures allows us to keep track of what is logically consistent and what is experimentally verifiable within a theory.
+This double structure allows us to keep track of what is logically consistent and what is experimentally verifiable within a theory.
 
 Finite time experimental verification also imposes a limit on how many statements one can verify.
-Given an arbitrarily long amount of time, a countable set of verifiable statement represents the maximum.
-Given a set $S \subseteq \vstmtSet$ of verifiable statements, we call a \textbf{basis} $\basis \subset S$ a subset from which all of $S$ can be generated through finite conjunction and countable disjunction.
+Given an arbitrarily long amount of time, a countable set of verifiable statements represents the maximum.
+Given a set $S \subseteq \vstmtSet$ of verifiable statements, we call a \textbf{basis} $\basis \subseteq S$ a subset from which all of $S$ can be generated through finite conjunction and countable disjunction.
 We define an \textbf{experimental domain} $\edomain_X \subseteq \vstmtSet$ a set of verifiable statements that includes the tautology and the contradiction, that is closed under finite conjunction and countable disjunction and that can be generated by a countable basis.
@@
 
-Within the theoretical domain, we define the  \textbf{possibilities of the domain} $X = \{ x \in \edomain_X \, | \, \forall \tstmt \in \tdomain_X $ either $x \narrower \tstmt$ or $x \ncomp \tstmt\}$ as those statements that, if true, will define the truth value for all other statements in the theoretical domain.
+Within the theoretical domain, we define the  \textbf{possibilities of the domain} $X = \{ x \in \tdomain_X \, | \, \forall \tstmt \in \tdomain_X $ either $x \narrower \tstmt$ or $x \ncomp \tstmt\}$ as those statements that, if true, will define the truth value for all other statements in the theoretical domain.
 These correspond to the cases that are experimentally distinguishable through the verifiable statements in the experimental domain.
 Conceptually, each possibility corresponds to a unique possible assignment for the domain.
-Mathematically, each possibility is a minterm of the basis (i.e.
-a conjunction where each element of the basis appears exactly once, either negated or not) that is not a contradiction.
+Mathematically, each possibility is a minterm of the basis (i.e.~a conjunction where each element of the basis appears exactly once, either negated or not) that is not a contradiction.
 
@@
 If a statement is verifiable, then $\ver(\tstmt) \equiv \tstmt$.
-Topologically, this corresponds to the interior of a Borel set and, indeed, an open set corresponds to its interior.
+Topologically, this corresponds to the interior of a Borel set and, indeed, an open set corresponds to its own interior.
 Conversely, we can define a falsifiable part $\fal(\tstmt)$ which, topologically, corresponds to the exterior of a Borel set.
-A statement that has nor interior or exterior, like the set of all rationals over a real line, corresponds to a statement for which the test never terminates: an undecidable statement.
+A statement that has neither interior nor exterior, like the set of all rationals over a real line, corresponds to a statement for which the test never terminates: an undecidable statement.
 One can proceed this way and re-understand all definitions and proofs.
@@
 The general result is that any set $X$ of experimentally distinguishable cases has to come with a $\mathsf{T}_0$ second countable topology that represents the verifiable statements and a Borel algebra that represents the theoretical statements.
-These structures are the foundation for most of the other mathematical structures used in science (e.g.
-differential topology, measure theory, probability theory, ...) and this framework provides a strong justification for them and a direct physical meaning.
+These structures are the foundation for most of the other mathematical structures used in science (e.g.~differential topology, measure theory, probability theory, ...) and this framework provides a strong justification for them and a direct physical meaning.
 It also gives us an understanding of exactly why those requirements are needed.
-A topological space that is not second countable, for example, would correspond to a set of possibility that cannot be, even in the limit, experimentally distinguished.
-A set with cardinality greater then that of the continuum, for example, does not allows a second countable topology and cannot correspond to a set of experimentally distinguishable cases: each element, in fact, must be identified with the output of a countable binary tests, which roughly corresponds to a real number in base two.
+A topological space that is not second countable, for example, would correspond to a set of possibilities that cannot be, even in the limit, experimentally distinguished.
+A set with cardinality greater than that of the continuum, for example, does not allow a second countable topology and cannot correspond to a set of experimentally distinguishable cases: each element, in fact, must be identified with the output of countable binary tests, which roughly corresponds to a real number in base two.
 A non-Borel set is also non-physical as it does not correspond to a statement for which we can construct a test, regardless of whether it terminates.
-In our view, much confusion is there in physics because we do not have such precise and methodical mapping between physical ideas and mathematical ones.
+In our view, confusion arises in physics because we do not have such precise and methodical mapping between physical ideas and mathematical ones.
 
@@
 This framework allows us to ask the question: when is a set of physically distinguishable cases identifiable with the real numbers?
-Under what conditions the points become isomorphic to those of a real line?
+Under what conditions do the points become homeomorphic to those of the real line?
 
@@
 
-If we concentrate on a single property of an object, we can select the verifiable statement that only pertain to that property.
+If we concentrate on a single property of an object, we can select the verifiable statements that only pertain to that property.
 These will form an experimental domain $\edomain_X$ whose possibilities $X$ will correspond to all possible property values.
 Recall that, given a totally ordered set $(\mathcal{Q}, \leq)$, the \textbf{order topology} is the one generated by sets of the form $$(a, \infty) = \{q \in \mathcal{Q} \, | \, a < q\} \;,\; (-\infty, b) = \{q \in \mathcal{Q} \, | \, q < b\}.$$
-The standard topologies for the integers and the reals are simply the order topologies with their respective ordering.
+The standard topologies for the integers and the reals are simply the order topologies under their respective ordering.
 We will say that an experimental domain is \textbf{naturally ordered} when the possibilities are totally ordered and the natural topology is the corresponding order topology.
@@
 For example, a mark on a ruler will also have an extent in space.
-Also note that, in general, the object being measured can also have an extent.
+Also note that, in general, the object being measured can have an extent.
 For example, an event may start before the tick of a clock and finish after.
@@
 Let us define a \textbf{reference} $\refStmt = ( \stmt[b], \stmt[o], \stmt[a] )$ as a tuple of three statements (the before/on/after statements) such that:
-\begin{enumerate}
-	\item we can verify whether the object is before or after the reference: $\stmt[b]$ and $\stmt[a]$ are verifiable statements
+\begin{itemize}
+	\item we can verify whether the object is before or after the reference: $\stmt[b], \stmt[a] \in \vstmtSet$
 	\item the reference itself must occupy some values : $\stmt[o] \nequiv \contradiction$
@@
 	\item if the object is before and after, it is also on the reference: $\stmt[b] \AND \stmt[a] \narrower \stmt[o]$
-\end{enumerate}
+\end{itemize}
 A set of references generates an experimental domain in the following way.
@@
 The \textbf{before domain} $\edomain_b$ is the domain generated only by the before statements $\basis_b$ and the \textbf{after domain} $\edomain_a$ is the domain generated only by the after statements $\basis_a$.
-The possibilities $X$ of the domain generated by $R$ consists of all the cases that are experimentally distinguishable given the references.
+The possibilities $X$ of the domain generated by $R$ consist of all the cases that are experimentally distinguishable given the references.
 
-We can see that if $(X, \leq)$ is a totally ordered set of mutually incompatible statements (i.e.
-no two can be true at the same time) whose union is a tautology (i.e.
-one of them must be true), we can create the collection of triples $R = \{ \bigOR\limits_{x_b < x} x_b , x , \bigOR\limits_{x_a > x} x_a \}_{x \in X}$.
+We can see that if $(X, \leq)$ is a totally ordered set of mutually incompatible statements (i.e.~no two can be true at the same time) whose union is a tautology (i.e.~one of them must be true), we can create the collection of triples $R = \{ \bigOR\limits_{x_b < x} x_b , x , \bigOR\limits_{x_a > x} x_a \}_{x \in X}$.
 If the before and after statements are verifiable, then the triples are references and the natural topology of the experimental domain they generate is exactly the order topology.
-Therefore, every naturally ordered experimental domains can be generated by a set of references.
-What we are looking is for the opposite conditions: what are the minimum requirements on a set of references such that they generate a naturally ordered experimental domain?
+Therefore, every naturally ordered experimental domain can be generated by a set of references.
+What we are looking is for the inverse conditions: what are the minimum requirements on a set of references such that they generate a naturally ordered experimental domain?
 
@@
 
-One of the main guiding insights is that, in a total order, $(a, +\infty) ^C = (-\infty, a]$.
-If $a \geq b$ then we have $(-\infty, b) \subseteq (-\infty, a]$ and if $a < b$ then $(-\infty, a) \subset (-\infty, b]$.
-This means that if $X$ is totally ordered, then all the sets of them form $(-\infty, x)$ and the complements of $(x, \infty)$ will be ordered by set inclusion.
+One of the main guiding insights is that, in a total order, $(x, +\infty) ^C = (-\infty, x]$.
+If $x \geq y$ then we have $(-\infty, y) \subset (-\infty, x]$ and if $x < y$ then $(-\infty, x] \subset (-\infty, y)$.
+This means that if $X$ is totally ordered, then all the sets of the form $(-\infty, x)$ and the complements of $(x, \infty)$ will be ordered by set inclusion.
 In terms of statement, this means that all the before statements $\basis_b$ and the negation of the after statements $\NOT(\basis_a)$ are ordered by narrowness.
-This, turns out, is a necessary condition.
+This, it turns out, is a necessary condition.
 
@@
 For a strict reference we then have $\stmt[b] \narrower \NOT \stmt[a]$.
-This is automatically achieved for discrete quantities, since object do not have an extent over them.
+This is automatically achieved for discrete quantities, since objects do not have an extent over them.
 For those that have an extent, for example when measuring the position of an object, this assumes that the extent of the object to be measured is very small compared to the extent of the reference so that the object can always be considered wholly before/on/after the reference.
@@
 Formally, one reference is before the other, noted $\refStmt_1 \less \refStmt_2$, if $\stmt[b]_1 \OR \stmt[o]_1 \ncomp \stmt[o]_2 \OR \stmt[a]_2$.
-If they do not overlap, they at least are such that we can imagine they have boundaries that come one after the other.
-Moreover, alignment implicitly requires that we can always tell references apart such that their relationship are always expected to be the same.
+If they do not overlap, they are such that we can at least imagine they have boundaries that come one after the other.
+Moreover, alignment implicitly requires that we can always tell references apart such that their relationship is always expected to be the same.
 If they jiggle around there can be no fixed, well-defined, alignment.
@@
 	\item find an intermediate one if they are not consecutive; that is, if $\refStmt[r]_1 < \refStmt[r]_2$ but $\refStmt[r]_2$ is not the immediate successor of $\refStmt[r]_1$ (i.e.
-$\NOT \stmt[a]_1 \snarrower \stmt[b]_2$), then we can find a strict reference $\refStmt_3$ aligned with $\edomain$ such that $\refStmt[r]_1 < \refStmt[r]_3 < \refStmt[r]_2$.
-	\item refine overlapping references if one is finer than the other; that is, if $\stmt[o]_2 \snarrower \stmt[o]_1$, we can find a strict reference $\refStmt_3$ aligned with $\edomain$ such that $\stmt[o]_3 \narrower \stmt[o]_1$ and either $\stmt[b]_3 \equiv \stmt[b]_1$ and $\refStmt_3 < \refStmt_2$ or $\stmt[a]_3 \equiv \stmt[a]_1$ and $\refStmt_2 < \refStmt_3$.
+$\NOT \stmt[a]_1 \snarrower \stmt[b]_2$), then we can find a strict reference $\refStmt_3$ such that $\refStmt[r]_1 < \refStmt[r]_3 < \refStmt[r]_2$.
+	\item refine overlapping references if one is finer than the other; that is, if $\stmt[o]_2 \snarrower \stmt[o]_1$, we can find a strict reference $\refStmt_3$ such that $\stmt[o]_3 \narrower \stmt[o]_1$ and either $\stmt[b]_3 \equiv \stmt[b]_1$ and $\refStmt_3 < \refStmt_2$ or $\stmt[a]_3 \equiv \stmt[a]_1$ and $\refStmt_2 < \refStmt_3$.
 \end{itemize}
-This conditions means that we are able to cover the whole space with references fine enough to resolve all possible overlaps.
-This is what allows us to pick out individual values as we can ultimately find a before statement that is broadest narrower statement of each negated after statement.
+These conditions means that we are able to cover the whole space with references fine enough to resolve all possible overlaps.
+This is what allows us to pick out individual values as we can ultimately find a before statement that is the broadest statement narrower than each negated after statement.
 
@@
 \begin{thrm}[Discrete ordering theorem]
-	An experimental domain is naturally ordered with a discrete quantity if and only if it is generated by a set of refinable aligned strict references with a sparse order (i.e.
-only finitely many references between two)
+	An experimental domain is naturally ordered by a discrete quantity if and only if it is generated by a set of refinable aligned strict references with a sparse order (i.e.~only finitely many references between any two)
 \end{thrm}
 \begin{thrm}[Continuous ordering theorem]
-	An experimental domain is naturally ordered by a continuous quantity if and only if it is generated by a set of refinable aligned strict references with a dense order (i.e.
-infinitely many references between two)
+	An experimental domain is naturally ordered by a continuous quantity if and only if it is generated by a set of refinable aligned strict references with a dense order (i.e.~infinitely many references between any two)
 \end{thrm}
 
-Note that the only extra condition to recover the continuous and the discrete orderings is the cardinality of references found between two.
-This may be unexpected for the continuous case, since the real numbers not only require the ordering be dense, but also complete and to have a countable dense subset.
-What happens is that the experimental domain automatically provides the other two conditions automatically.
+Note that the only extra condition to recover the continuous and the discrete orderings is the cardinality of references found between any two.
+This may be unexpected for the continuous case, since the real numbers not only require the ordering to be dense, but also complete and to have a countable dense subset.
+What happens is that the experimental domain provides the other two conditions automatically.
 The closure of the domain under countable disjunction requires the ordering to be complete.
@@
 
-Now that we have understood what are the requirements to experimentally define a continuous quantity, we can ask the following question: as we increase spatial and temporal resolution and we approach Plank scale, can we still satisfy those conditions?
+Now that we have understood what the requirements to experimentally define a continuous quantity are, we can ask the following question: as we increase spatial and temporal resolution and we approach Planck scale, can we still satisfy those conditions?
 Or can they only be satisfied as an approximation at large enough scale?
 
-At this point, we do not aim lay out a convincing argument for or against.
-Rather, we want to show what are the main points, both conceptually and technically, that one may raise so that one can understand how to construct his own argument.
+At this point, we do not aim to lay out a convincing argument for or against.
+Rather, we want to show what the main points are, both conceptually and technically, so that one can understand how to construct his own argument.
 We believe this to be more interesting because, once one decides how and why those assumptions fail, he is already on the path to understand what alternative he wants to explore.
@@
 One of the requirements is that references are strict, that is the before/on/after cases are mutually exclusive.
-When measuring position in space, where finite objects have a finite extent, those cases are not in general mutually exclusive because of overlaps.
+When measuring position in space, where objects have a finite extent, those cases are not in general mutually exclusive because of overlaps.
 We can ignore the problem by assuming that the object whose position we want to measure is much smaller than the reference, in which case we can assume that the object is wholly before/on/after the reference.
 But in order to derive continuous quantities, we also assume that we can take finer and finer references.
-If the objects we use as references of position are the same objects of which we want to measure the position we logically come at an impasse.
-This is going to be a problem for particles (i.e.
-fields excitation) as well since the overlap can be in terms of their wave-packet.
-In these cases, it is not clear one can make the claim that a particle will always be wholly before/on/after the other.
+If the objects we use as references of position are the same objects of which we want to measure the position we logically come to an impasse.
+This is going to be a problem for particles (i.e.~field excitations) as well since the overlap can be in terms of their wave-function.
+In these cases, it is not clear one can make the claim that one particle will always be wholly before/on/after another.
 
-Another requirement is that references are aligned, that is we can establish clear before/after relationship between the edges of the references.
-One issue here is that we always need to be able to distinguish between two references such that fixed ordering relationship can be established.
+Another requirement is that references are aligned, that is we can establish a clear before/after relationship between the edges of the references.
+One issue here is that we always need to be able to distinguish between two references such that a fixed ordering relationship can be established.
 If the finest references are individual particles, and these are indistinguishable from each other, it is not clear how this is going to work.
 Note that we cannot use position itself to distinguish them because we run into circular arguments: we want to use the reference to distinguish position not the other way around.
-Another problem is that we cannot talk about a substructure for individual particles, therefore it is not clear whether the edges (i.e.
-where the particle ``starts") can be well defined if particles are used as references.
+Another problem is that we cannot talk about a substructure for individual particles, therefore it is not clear whether the edges (i.e.~where the particle ``starts") can be well defined if particles are used as references.
 
-Another requirement is that references are refinable, that is we can always find finer references that resolve the overlap or be placed in gaps.
+Another requirement is that references are refinable, that is we can always find finer references that resolve the overlap or can be placed in gaps.
 While it is not necessary to have them all in the same setup as we can, in principle, repeat the experiment with different references in place to get different information, we still need to be able to construct them.
-While it is generally accept that there is a limit to how small a wave-packet can be made, meaning how small the extent of an object (or a reference) can be, there is no such limit between the distance between two objects.
+While it is generally accepted that there is a limit to how small a wave-packet can be made, meaning how small the extent of an object (or a reference) can be, there is no such limit between the distance between two objects.
 Therefore we can imagine creating situations in which two references are consecutive and yet no other reference can fit in between; or two references overlap yet no other reference can refine the overlap.
 
-The punchline here is that if a single argument along these lines is accepted, the result is that we lose not just the real numbers, we lose ordering itself.
+The punchline here is that if any single argument along these lines is accepted, the result is that we lose not just the real numbers, we lose ordering itself.
 What would happen is that, at short scales, ordering becomes fuzzy.
@@
 With a loss of ordering, there is not a clear way to keep a sharp notion of distance and all other geometrical ideas.
-What we would have in its place would depend on how the structure breaks down and what new assumptions can be taken.
+What we would have in its place would depend on how the structure breaks down and what new assumptions can be made.
 
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 \bibliography{bibliography}

### 2020-03-31 Articles/2019-QuantumComposition/Source/MathCheck.tex
Version put in the archive

%Dovremmo forse aggiungere due parole sulle C* algebras: non sono stateintrodotte proprio per evitare il tensor product in infinitedimensional systems?
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% The four postulates of quantum mechanics are three
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[aps,prl,amsmath,amssymb,twocolumn,nofootinbib]{revtex4}
%\documentclass[twocolumn,aps,showpacs,prl,groupedaddress]{revtex4}
%\usepackage{amssymb,stackengine}
\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epsfig,amssymb,amsmath,amsthm}
%\usepackage[active]{srcltx}
%\usepackage[hypertex,linkcolor=red]{hyperref}

\usepackage{color}
\usepackage{cancel}
\usepackage{tikz-cd}

\theoremstyle{plain}
\newtheorem{thrm}{Theorem}[section]
\newtheorem{form}[thrm]{Formalization}
\newtheorem{prop}[thrm]{Proposition}
\newtheorem{lem}[thrm]{Lemma}

\theoremstyle{definition}
\newtheorem{defn}[thrm]{Definition}
\newtheorem{post}{Postulate}[]
\renewcommand*{\thepost}{(\alph{post})}

\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\newcommand{\blue}{\color{blue}}  %NON LINEAR
\newcommand{\red}{\color{red}}
\newcommand{\cyan}{\color{cyan}}
\newcommand{\green}{\color{green}} %THEORY quantum state estimation
\newcommand{\yellow}{\color{yellow}} %THEORY quantum channel estimation

\DeclareMathOperator{\spn}{span}

\newcommand{\pj}[1] {\underbar{$#1$}}
%\newcommand{\pj}[1] {\overline{#1}}

\def\>{\rangle}
\def\<{\langle}
\def\ca{_{\cal A}}
\def\cb{_{\cal B}}
\def\cc{_{\cal C}}
%\def\comment#1{}
\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
\def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
\def\labell#1{\label{#1}}
%\def\labell#1{\label{#1}{\mbox{{\tiny #1}}}}
%\def\section#1{{\par\em #1:--- }}
\def\togli#1{}
\def\sh{\mbox{sh}}
\def\iden{\openone}
\begin{document}

	\fbox{{\scriptsize Internal report. \today}}\title{The four postulates of quantum mechanics are three}
	\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^2$ and Christine A. Aidala$^1$
	}\affiliation{\vbox{1.~Physics Department, University of Michigan, 450 Church Street,
			Ann Arbor, MI 48109-1040,
			United States}\\
		\vbox{2.~Dip.~Fisica and INFN Sez.\ Pavia, University
			of Pavia, via Bassi 6, I-27100 Pavia, Italy}}
	\begin{abstract}
		The tensor product postulate of quantum mechanics states that the
		Hilbert space of a composite system is the tensor product of the
		components' Hilbert spaces. All current formalizations of quantum
		mechanics that do not contain this postulate contain some equivalent
		postulate or assumption (sometimes hidden). Here we give a natural
		definition of composite system as a set containing the component
		systems and show how one can logically derive the tensor product
		rule from the state postulate and from the measurement postulate. In
		other words, our paper reduces by one the number of postulates
		necessary to quantum mechanics.
	\end{abstract}
	\pacs{}
	% Measurement theory (quantum mechanics), 03.65.Ta Mechanics quantum,
	% 03.65.-w noise quantum, 42.50.Lc quantum information, 03.67.Ac
	% Quantum information, 03.67.-a Quantum fluctuations, 42.50.Lc quantum
	% mechanics, 03.65.Ta quantum optics, 42.50.Gy
	\maketitle

	The tensor product postulate does not appear in all axiomatizations of
	quantum mechanics: it has even been called ``postulate 0'' in some
	literature \cite{zurek}. A widespread belief is that it is a direct
	consequence of the superposition principle, and it is hence not a
	necessary axiom. {\em This belief is mistaken}: the superposition
	principle is encoded into the quantum axioms by requiring that the
	state space is a {\em linear} vector space. This is, by itself,
	insufficient to single out the tensor product, as other linear
	products of linear spaces exist, such as the direct product, the
	topological product or the direct sum of vector spaces, which are used
	in classical mechanics to combine state spaces of linear systems.
	This belief may have arisen from the seminal book of Dirac
	\cite{diracbook}, who introduces tensor products (Chap.~20) by
	appealing to linearity. However, he adds the seemingly innocuous
	request that the product among spaces be distributive (rather,
	bilinear), which is equivalent to postulating tensor products (or
	linear functions of them). This is not an innocuous request. For
	example it does not hold where the composite vector space of two
	linear spaces is described by the direct product, e.g.~in classical
	mechanics, for two strings of a guitar: it is not distributive.
	[General classical systems, not only linear ones, are also composed
	through the direct product.] Of course, Dirac is not constructing an axiomatic
	formulation, so his `sleight of hand' can be forgiven. In contrast, von
	Neumann (\cite{vonneumannbook} Chap.~VI.2, also \cite{jauch})
	introduces tensor products by noticing that this is a natural choice
	in the position representation of wave mechanics (where they were
	introduced in \cite{weyl,epr}), and then {\em explicitly postulates}
	them in general: ``This rule of transformation is correct in any case
	for the coordinate and momentum operators [...] and it conforms with
	the [observable axiom and its linearity principles], we therefore
	postulate them generally.''  \cite{vonneumannbook}.  More mathematical
	or conceptually-oriented modern formulations
	(e.g.~\cite{ozawa,masanes,nielsenchuang}) introduce this postulate explicitly.
	An interesting alternative is provided in
	\cite{ballentinebook,ballentinepaper}: after introducing tensor
	products, Ballentine verifies a posteriori that they give the correct
	laws of composition of probabilities. Similarly, Peres uses
	relativistic locality \cite{peres}. While these procedures seemingly
	bypass the need to postulate the tensor product, they do not guarantee
	that this is the {\em only} possible way of introducing composite
	systems in quantum mechanics. In the framework of quantum logic,
	tensor products arise from some additional conditions \cite{matolcsi}
	which (in contrast to what is done here) are not connected to the
	other postulates. A similar approach was followed in \cite{aerts}
	where tensor products were obtained by specifying some additional
	physical requirements. In quantum field theory one tends to avoid
	problems connected with tensor products of infinite dimensional spaces
	by focusing on algebraic commutation structures, e.g.~\cite{giddins}.
	In particular, the recent MIP*=RE result \cite{mipre} implies that, in
	infinite dimensions, the tensor product is strictly less
	computationally powerful than the commutation structures, emphasizing
	the difference among these two structures, at least for the
	infinite-dimensional case. We will consider the non-relativistic
	setting here.

	In this paper we derive the tensor product postulate (which, hence,
	loses its status of postulate) from two other postulates of quantum
	mechanics: the state postulate and the measurement postulate. We start
	from the natural definition of a composite system as the set of two
	(or more) quantum systems, in the sense that the composite system is
	made of system $A$ {\em and} (joined with) system $B$, namely a set
	whose elements are the two systems and nothing else. We will focus on
	kinematically-independent systems, namely no superselection rules or
	other restrictions to the state space are present: it is possible to
	prepare each subsystem of a composite system in a state that is
	independent of the other systems (preparation independence).  This is
	the only case in which the tensor product can be properly employed
	\cite{susskind,zanardi,zanardilloyd}: the Hilbert space of composite
	systems that have restrictions is {\em not} the tensor product of the
	component spaces, but a subspace of it (e.g.~the anti-symmetric
	subspace for fermions). Typically this is ignored in the literature,
	since the tensor product formalism is very convenient and is often
	used also in these cases. Moreover, in accordance with the measurement
	postulate, we require that the probability distribution of the
	measurement outcomes of one system is independent of the other systems' (statistical
	independence). As detailed below, statistical independence is not an
	additional requirement: it is already contained in the measurement
	postulate.

	From the above definition of composite system, it follows that there
	must be a map $M$ that connects the states of the subsystems to the
	states of the composite system. A quantum state is a ray in Hilbert
	space, namely a set of vectors. The map $M$ on states 
	corresponds to a map $m$ on vectors. We then show that preparation
	independence and statistical independence imply three conditions on
	the map $m$: (H1)~totality: the map is defined on all states of the
	subsystems; (H2)~bilinearity: the map is bilinear thanks to the
	fundamental theorem of projective geometry; (H3)~span surjectivity:
	the span of the image of map coincides with the full composite Hilbert space.
	We then prove that, if these three conditions H1, H2 and H3 hold, then
	the map $m$ is the tensor product, namely the Hilbert space of the
	composite system is a tensor product of the components: the tensor
	product ``postulate'', which hence loses its status of a postulate. An
	overview of all these logical implications is given in
	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
	argument. Refer to the supplementary material for a more rigorous proof of the \emph{same} argument.

	\begin{figure}[h]
		\includegraphics[width=\linewidth, trim={0.2in 5.2in 2.5in 0.6in}, clip=true]{fig.eps}
		\caption{Schematic depiction of the logical implications used here. FTPG stands for ``Fundamental Theorem of Projective Geometry''
			\label{f:fig}}\end{figure}
	\togli{
		\begin{widetext}\begin{verbatim}
			* P1: states and observable postulate.
			* P2: Born rule (measurement postulate). 
			* P1 -> systems can be prepared unconditionally in any state.
			* P2 -> the outcome probabilities depend only on the scalar
			product.
			* 2. Def: Composite system definition: A composite system is a collection of
			(all compatible state give a 
			preparation) and only of (all composite preparations give non-trivial
			measurement on subsystems) the subsystems 
			* P1+ Def -> 1. Preparation independence (all state pairs are compatible)
			* P2+ Def -> 3. Span surjectivity (I.4=H3) (all C is spanned by AxB) 
			* 3 -> 6. Basis carries over I.7
			* 1 + 2 -> 4. Injectivity (M total function) I.5=H1
			* 2 + PI -> 5. Statistical independence I.6 (probability on compatible
			pairs where one subsystem does not change) 
			* 5 -> 7. I.8=H2: scalar product carries over (fundamental theorem of
			projective geometry).
			* 7 + 6 + 4 (i.e. H1+H2+H3) -> 8. m is tensor product   
			\end{verbatim}\end{widetext}
	}

	We start by noting that the tensor product is uniquely
	characterized, up to isomorphism, by a universal property regarding
	bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
	tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
	map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the property
	than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ factors
	through $T$ uniquely.  This means that there exists a {\em unique}
	$\hat m$, dependent on $m$, such that $\hat m \circ T=m$.  In other
	words, the following diagram commutes:
	\begin{center}
		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{m}"] \\
			& \mathcal{C}
		\end{tikzcd}
	\end{center}

	We use the axiomatization of quantum mechanics based on the following
	postulates (e.g.~\cite{ozawa,nielsenchuang}): (a)~The state of a
	system is described by a ray $\pj{\psi}$ corresponding to a set of
	non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
	system's observable properties are described by self-adjoint operators
	acting on that space; (b)~The probability that a measurement of a
	property $X$, described by the operator with spectral decomposition
	$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
	$x$ given that the system is in state $\pj{\psi}$ is
	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2$ (Born rule). (c)~The state
	space of a composite system is given by the tensor product of the
	spaces of the component systems; (d)~The time evolution of an isolated
	system is described by a unitary operator acting on a vector
	representing the system state, $|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or,
	equivalently, by the Schr\"odinger equation. The rest of quantum
	theory can be derived from these axioms. While some axiomatizations
	introduce further postulates, we will be using only (a) and (b) to
	derive (c), so the above are sufficient for our aims.

	\togli{This axiomatization implicitly contains a definition of ``quantum
		system'' which is crucial for what follows, so we need to clarify the
		assumptions that it contains. We will use the following definition for
		a quantum system\togli{$\stackon[1pt]={\mbox{\tiny
					def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree of
			freedom with $d$ (possibly discrete, or continuous, infinite)
			mutually exclusive (commuting) values for each of its properties.
			Its mathematical description is through a Hilbert space of dimension
			$d$ which contains all the states that describe the values of its
			possible properties. In accordance with the postulate (a), these
			values correspond to a basis of the space, given by the eigenvectors
			of the observable corresponding to that property''}. \commentg{We may have to revise to be more clear. Where is this used?} }
	As stated
	above, we will limit ourselves to kinematically-independent systems,
	where all state vectors $|\psi\>$ in the system's Hilbert space $\cal H$
	describe a valid state, {\em unconditioned on anything else}. In
	particular, we note that restrictions due to superselection rules
	arise either from practical (not fundamental) limitations on the
	actions of the experimenter \cite{susskind,zanardi,zanardilloyd} or
	from the use of ill-defined quantum systems. For example, in
	situations where indistinguishability plays a role (e.g.~in QFT), one
	cannot consider an electron as a quantum system: in this case the
	quantum system is the field. The electron is an excitation of the
	field, {\em not} a (sub)system.

	This implies that the ``system'' is kinematically independent from
	anything else.  We call this condition ``preparation independence''.
	[We emphasize that the kinematic independence is inequivalent to
	dynamical independence (or isolation).  Indeed if two systems
	interact, their interaction may lead to dynamical restrictions in the
	state spaces. We will not consider dynamical evolution in this paper,
	which is contained in postulate (d).]

	The definition of a composite system as containing {\em only} the
	collection of the subsystems means that any preparation of both
	subsystems independently must correspond to the preparation of the
	composite system. Since states are defined by postulate (a) as rays in
	the respective Hilbert spaces, there must exist a map $M : \pj{\cal A}
	\times \pj{\cal B } \to \pj{\cal C}$ that takes a pair of states for
	the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$ represent the
	projective spaces and the Cartesian product is the set of all possible
	pairs) and returns a state in the projective space $\pj{\cal C}$ for
	the composite. The map $M$ that acts on rays corresponds to a map $m:{\cal A}\times{\cal B}\to{\cal C}$ that
	acts on vectors in the Hilbert spaces $\cal A$, $\cal B$ and $\cal C$.
	Namely, $\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign
	indicates the elements in the projective space. We will prove that the
	map $m$ is the tensor product.

	The map $M$ must be injective: as said above, different states of the
	subsystems must correspond, by definition of composite system, to
	different states of the composite. Moreover, preparation independence
	implies that $M$, and hence $m$, must be total maps: each subsystem of
	the composite system can be independently prepared and gives rise to a
	state of the composite (condition H1).  H1 is not sufficient
	to identify the tensor product: by itself it does not even guarantee
	that the map $m$ is linear.

	Postulate (b) contains the connection between quantum mechanics
	and probability theory. It must then implicitly contain the
	axiomatization of probability, e.g.~see
	\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
	probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
	the joint probability of events $a$ and $b$ given $z$ is
	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Then the events $a$ and $b$ are
	independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$ (this
	is the {\em definition} of independent events). Since the Born
	probability formula in postulate (b) contains only quantities of the
	system (the state and the observable's eigenstates {\em of the
		system}), it implies the ``statistical independence'' between
	different systems if they are prepared independently. Namely, the
	outcome probability of one is independent of any properties of any
	other in the absence of any dynamical coupling and supposing
	independent preparation. [Of course, the dynamics may couple the state
	of the system to other systems so that the outcomes may depend on what
	happens to other systems.  Analogously, one may prepare the system in
	a way that depends on other systems. But we consider independent
	preparations and do not consider dynamics, as it is sufficient for our
	aims.] We can then formalize ``statistical independence'' of
	independently prepared systems by saying that their Born rule
	satisfies:
	\begin{eqnarray}
	p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
		B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
		A})\:p(b|\psi_{\cal B})
	\labell{bindip}\;,
	\end{eqnarray}
	where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
	preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
	$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
	B$ respectively. The first equality in \eqref{bindip} embodies the
	definition of statistical independence, the second follows from the
	form of the Born rule for independently prepared systems (the
	probability depends only on the state and observables of each system
	on its own).

	A particular case is when one system is
	prepared in a fixed eigenstate $|b\>$ of the observable measured on it,
	$p(a\wedge b|\psi\wedge b)=p(a|\psi)\:p(b|b)=p(a|\psi)$, where the last equality follows from the Born rule since $\<b|b\>=1$. 
	If we define $M_b(\pj{a}) = M(\pj{a},\pj{b})$, substituting the values of the
	probabilities from the Born rule, we have:
	\begin{align} &\Big|\Big\<M\left(\pj{a},\pj{b}\right)\Big|M\left(\pj{\psi},\pj{b}\right)\Big\>_{\cal C}\Big|^2
	=\Big|\Big\<M_b\left(\pj{a}\right)\Big|M_b\left(\pj{\psi}\right)\Big\>_{\cal C}\Big|^2
	\nonumber \\&
	=\left|\<a|\psi\>_{\cal A}\right|^2
	\labell{questa},
	\end{align}
	where the first and second terms contain the scalar product in the composite
	space $\cal C$. [This is not a new assumption: it follows from the
	measurement postulate (b) for the composite system.] This means that,
	when one subsystem is prepared in an eigenstate of what is measured
	there, the state space of the other is mapped preserving the square of
	the inner product.

	Since the square of the inner product is preserved, orthogonality and
	the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
	colinear transformation by definition. In this case, the fundamental
	theorem of projective geometry \cite{fun} applies, which tells us that
	a unique semi-linear map $m_b$ that acts on the vectors exists in accordance with $M_b$.
	Moreover, conservation of probability further constrains it to be
	either linear or antilinear. This tells us that the corresponding $m$
	is either linear or antilinear in the first argument. Namely, if
	\eqref{questa} holds, then
	\begin{align}
	\<a|\psi\>=\<m(a,b)|m(\psi,b)\>\labell{h2}\;
	\\\mbox{ or }
	\<a|\psi\>=\<m(\psi,b)|m(a,b)\> \labell{h2b}.
	\end{align}
	We can ignore the antilinear case \eqref{h2b}, which simply
	corresponds to the map $m$ mapping kets of one or both subsystems into
	bras before outputting the composite system (all these maps are
	trivially physically equivalent, since the dual-space description of a
	quantum system through bras is equivalent to the description using
	kets). We can now repeat the same analysis for the second argument of
	$m$ to conclude that it is a bilinear map, condition (H2).

	The last condition (H3) follows directly from the definition of a
	composite system. Since it is composed {\em only} of the component
	systems, for any state $c$ of the composite system, we must find at least one pair $(a, b)$ such that $p(a\wedge b | c)\neq 0$. It follows that the map $m$ is span-surjective: namely the
	span of the map applied to all states in the component systems spans
	the composite system state space. In other words, the composite does
	not contain states that are totally independent of (i.e.~orthogonal
	to) the states of the components.

	We have obtained the conditions H1, H2 and H3 from the state postulate
	(a), the measurement postulate (b) and the definition of composite
	systems. We now prove that these three conditions imply that the (up
	to now unspecified) composition rule $m$ is the tensor product. More
	precisely, given a total, span-surjective, bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ that preserves the square of the inner
	product, we find that $\cal C $ is equivalent to
	$\cal A \otimes \cal B $ and that $m=\otimes$.

	Proof. Step 1: the bases of the component systems are mapped to a
	basis of the composite system. Because of totality property (H1) and
	because the square of the inner product is preserved, we can conclude
	that, given two orthonormal bases $\{|a_i\>\}\in{\cal A}$ and
	$\{|b_j\>\}\in{\cal B}$,
	$|\<m(a_i,b_j)|m(a_k,b_\ell)\>|^2=\delta_{ik}\delta_{j\ell}$, namely
	$\{|m(a_i,b_j)\>\}$ is an orthonormal set in $\cal C$.  Moreover, the
	surjectivity property (H3) guarantees that in $\cal C$ no vectors are
	orthogonal to this set. This implies that it is a basis for $\cal C$.

	Step 2: Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is
	a bilinear operator (property H2), thanks to the universal property of
	the tensor product we can find a unique linear operator $\hat{m} :
	\mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that $m(a, b) =
	\hat m(a \otimes b)$. The set $\{ \hat m(a_i\otimes b_j)$ with
	$|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and ${\cal B}\}$
	forms a basis for $\cal C$, since $\hat m(a_i\otimes b_j)=m(a_i,b_j)$
	and we have shown above that the latter is a basis.  Thus,
	\begin{align} 
	&\<\hat m(a_i\otimes b_j)|\hat m(a_k\otimes b_\ell)
	\>_{\mathcal{C}}=\<m(a_i,
	b_j)|m(a_k,b_\ell)\>_\mathcal{C} \nonumber\\& =
	\delta_{ik}\delta_{j\ell}
	= \<a_i\otimes
	b_j| a_k \otimes b_\ell\>_{\otimes},
	\labell{ecco}\; 
	\end{align}
	where we used the orthonormality of the bases and the fact that
	$|a_i\otimes b_j\>$ is a basis of the tensor product space ${\cal
		A}\otimes{\cal B}$. The function $\hat{m}$, then, is an isomorphism
	over all elements of the basis $|a_i\otimes b_j\>$. It is then an
	isomorphism between $\cal C$ and ${\cal A}\otimes{\cal B}$. Namely,
	$|m(a_i,b_j)\>_{\cal C} = |\hat{m}(a_i\otimes b_j)\>_{\cal C} \cong
	|a_i\otimes b_j\>_\otimes$, where the first two vectors (in $\cal C$)
	are isomorphic to the last vector (in the tensor product space ${\cal
		A}\otimes{\cal B}$). Namely, $\hat{m}$ maps bijectively to the
	tensor product, which implies the isomorphism ${\cal C}\cong{\cal
		A}\otimes{\cal B}$, i.e.~their equivalence through the map
	$\hat{m}$. Moreover, since ${\cal C}\cong{\cal A}\otimes{\cal B}$,
	then the map $m : \cal A \times \cal B \to \cal C$ is equivalent to
	the map $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$,
	i.e.~the maps $m$ and $\otimes$ are also equivalent through $\hat
	m$.$\square$

	\vskip 1\baselineskip A few comments on the proof: it is based on the
	universal property of the tensor product, which uniquely characterizes
	it. In Step 1 we showed that the bilinear map $m$ maps subsystems'
	bases into the composite system basis. We also know that there exists a
	tensor product map $ {T}=\otimes$ that can compose the vectors in $\cal A$
	and $\cal B$. In Step 2 we use the universal property: since $m$ is a
	bilinear map, we are assured that there exists a unique $\hat m$ such
	that $\hat m \circ T=m$. Since we show that $\hat m$ is an
	isomorphism, then $\hat{m}$ bijectively maps vectors in $\cal C$
	onto vectors in the tensor product space. Namely $m={T}=\otimes$.

	We conclude with some general comments. The tensor product structure
	of quantum systems is not absolute, but depends on the observables
	that are accessible \cite{zanardi,zanardilloyd}. This is due to the
	fact that an agent that has access to a set of observables will define
	quantum systems differently from an agent that has access to a
	different set of observables. Where one agent sees a single system, an
	agent that has access to less refined observables (and is then limited
	by some superselection rules) can consider the same system as composed
	of multiple subsystems. A typical example \cite{tellerbook} comes from
	quantum field theory. It is customary in basically all quantum optics
	literature to treat different modes of the radiation field (e.g.~the
	output of two lasers) as independent systems composed through the
	tensor product.  Clearly the electromagnetic field is a single system
	and an agent who is able to access an optical mode that is a linear
	combination of the two will give a quantum description for it that
	cannot easily accommodate tensor products. Similarly, an agent can
	consider two electrons as two systems, joined with the tensor product,
	whenever they are distinguishable for all practical purposes (e.g.~the
	electrons are in widely separated physical locations). Yet, in
	principle, electrons are just excitations of a field, and the `true'
	quantum system is the field, not the single electrons
	\cite{teller,tellerbook}.  So, in quantum field theory, the quantum
	systems that should be joined through tensor products are the
	different fields and {\em not} the particles, which are just
	excitations (states) of the fields. In the words of Teller
	(\cite{tellerbook}, pg.22), tensor products can be safely used only if
	there is a ``primitive thisness'', which is captured in the definition
	of system.\togli{\comment{forse e' da spiegare meglio}}

	\togli{\comment{Commento finale da aggiungere?  E' interessante perche' tutto
			cio' ci dice che non otteniamo necessariamente il prodotto tensore,
			ma il prodotto tensore modulo una fase locale che e' fisicamente
			irrilevante. Viene da chiedersi se c'e' una qualche notazione
			(evidentemente piu' generale di quella di Dirac) che ci permetta di
			eliminare questa ambiguita' fisicamente irrilevante... La notazione
			di Dirac gia' elimina la necessita' di avere una rappresentazione
			per descrivere gli stati. Pero' evidentemente non elimina del tutto
			la ambiguita' di rappresentazione perche' un cambio di coordinate
			sia sugli stati che sugli operatori mi lascia invariata la fisica,
			ma non lascia invariata la notazione di Dirac...  Probabilmente la
			notazione che elimina l'ambiguita' e' quella che utilizza le matrici
			densita' invece dei vettori di stato (vedi Ozawa \cite{ozawa} e
			Holevo): le matrici densita' normalizzate sono phase-independent:
			$\rho=|\psi\>\<\psi|$. Attenzione: l'interpretazione di una matrice
			mista $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$ come ``il sistema e' nello
			stato $|\psi_i\>$ con probabilita' $p_i$'' e' una CONSEGUENZA della
			regola di Born, quindi nella formalizzazione del postulato degli
			stati in termini di matrici densita', questa interpretazione non
			puo' apparire perche' e' una conseguenza di un altro postulato.}}

	It has been pointed out before that the quantum postulates are
	redundant: in \cite{masanes} it was shown that the measurement
	postulate (b) can be derived from the others (a), (c), (d). Here
	instead we have shown how the tensor product postulate can be
	logically derived from the state postulate (a), the measurement
	postulate (b) and a reasonable definition of independent systems, and
	we have described the logical relations among them.  Of course, we do
	not claim that this is the {\em only} way to obtain the tensor product
	postulate from the others.

	L.M. acknowledges useful discussions with M.~Ozawa, P.~Zanardi,
	S.~Lloyd, D.~Zeh, G.~Auletta, A.~Aldeni. We acknowledge funding from
	the Attract project through the Eu Horizon 2020 research and
	innovation programme under grant agreement No 777222.
	\begin{references}
		\bibitem{zurek} W.H. Zurek, Quantum Darwinism, Nature Phys. {\bf 5},
		181 (2009).
		\bibitem{diracbook}P.A.M. Dirac, The principles of quantum mechanics,
		(Clarendon Press, Oxford, 1966).
		\bibitem{vonneumannbook}J. von Neumann, Mathematical Foundations of
		Quantum Mechanics (Princeton Univ.  Press, 1955).
		\bibitem{jauch}J.M. Jauch, Foundations of quantum mechanics
		(Addison-Welsey, 1968), pg.~176.
		\bibitem{weyl} H. Weyl, Gruppentheorie und Quantenmechanik (Hirzel,
		Leipzig, 1928); translated by H. P. Robertson, The Theory of Groups
		and Quantum Mechanics (Methuen, London, 1931); reprinted by Dover,
		p. 91.
		\bibitem{epr}A. Einstein, B. Podolsky, N. Rosen, Can
		quantum-mechanical description of physical reality be considered
		complete?, Phys. Rev. {\bf 47}, 777 (1935).
		\bibitem{ozawa}M. Ozawa, {Uncertainty relations for noise and
			disturbance in generalized quantum measurements}, Ann. Phys.  {\bf
			311}, 350 (2004).
		\bibitem{masanes}L. Masanes, T.D. Galley, M.P. M\" uller, The
		measurement postulates of quantum mechanics are operationally
		redundant, Nat. Commun. {\bf 10}, 1361 (2019).
		\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum Computation
		and Quantum Information (Cambridge University Press, Cambridge,
		2000).
		\bibitem{ballentinebook}L.E. Ballentine, Quantum Mechanics, a modern
		development (World Scientific, 2014).
		\bibitem{ballentinepaper}L.E. Ballentine, Probability theory in
		quantum mechanics, Am. J. Phys. {\bf 54}, 883 (1986).
		\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
		press, 1961).
		\bibitem{peres}A.~Peres, Classical interventions in quantum systems.
		II. Relativistic invariance, Phys. Rev. A {\bf 61}, 022117 (2000).
		\bibitem{matolcsi} T. Matolcsi, Tensor product of Hilbert lattices and
		free orthodistributive product of orthomodular lattices, Acta Sci.
		Math. (Szeged), {\bf 37}, 263 (1975).
		\bibitem{aerts} D. Aerts, I. Daubechies, Physical justification for
		using the tensor product to describe two quantum systems as one
		joint system, Helv. Phys. Acta {\bf 51}, 661 (1979).
		\bibitem{giddins}S.B.~Giddings, Hilbert space structure in quantum
		gravity: an algebraic perspective. J. High Energ. Phys. 2015, 1
		(2015).% https://doi.org/10.1007/JHEP12(2015)099
		\bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
		rule, Phys. Rev. {\bf 155}, 1428 (1967).
		\bibitem{zanardi}P. Zanardi, Virtual Quantum Subsystems, Phys. Rev.
		Lett {\bf 87}, 077901 (2001).
		\bibitem{zanardilloyd} P. Zanardi, D.A. Lidar, S. Lloyd, Quantum
		Tensor Product Structures are Observable Induced, Phys. Rev. Lett.
		{\bf 92},060402 (2004).
		\bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
		Field Theory (Princeton Univ. Press, 1997).  
		\bibitem{teller}M. Redhead, P. Teller, Particles, Particle Labels, and
		Quanta: The Toll of Unacknowledged Metaphysics, Found. Phys. {\bf
			21}, 43 (1991).
		\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
		\bibitem{holevo}A. Holevo, Probabilistic and statistical aspects of
		quantum theory, (North Holland, 1982).
	\end{references}

	\vskip 5\baselineskip

	\section{Appendix: mathematical formulation}\label{app}
	\setcounter{section}{1}

	The following section needs to distinguish the projective space from
	the Hilbert space itself. As this is seldom done in quantum mechanics,
	we review some concepts in that context and introduce the notation
	that we will be using. If $X$ is a Hilbert space, we denote $\pj{X}$
	the projective space. The projective space is mathematically
	constructed from the Hilbert space by removing the origin and
	quotienting by the equivalence relationship $v \sim \lambda v$, $v\in
	X$ and $\lambda\in\mathbb{C}$. A quantum state is a point in
	projective space. Each point of the projective space is called a ray,
	because for a real vector space to would correspond to a line going through
	the origin, with the origin removed. As we are in a complex space, the
	ray should be thought as a complex plane without the origin, which is
	the space of the vectors reachable from a fixed one through
	multiplication by a complex number. It can also be thought as a
	subspace of dimension one.

	Given a vector $v \in X$, we will denote $\pj{v}$ the ray in the
	projective space corresponding to $v$. Note that $\pj{v}$ denotes a
	quantum state, without having picked a modulus or phase. Given two or
	more vectors $v_1, ..., v_n \in X$, the subspace of $X$ they span
	(i.e.~all the vectors reached by linear combinations) is noted by
	$Sp(v_1, ..., v_n)$. Note that this subspace will correspond to a set
	of rays in the projective space, which we note as $\pj{Sp(v_1, ...,
		v_n)}$. Given $v,w \in X$, we can write $P(v|w) = \frac{|\< v | w
		\>|^2}{\< v | v \>\< w | w \>}$ which corresponds to the probability
	of observing $v$ given $w$ was prepared. Note that $P(v|w) = P(\lambda
	v| \mu w)$, with non-null $\lambda,\mu\in{\mathbb C}$, and therefore
	one can write $P(\pj{v} | \pj{w})\equiv P(v|w)$ as a function of the
	rays.

	\begin{post}\label{post_state}
		The state of a quantum system is described by a ray $\pj{\psi} = \{
		\alpha |\psi\> \, |$ non-null $\alpha \in
		\mathbb{C},|\psi\>\in\mathcal{H}\}$ in a complex Hilbert space
		$\mathcal{H}$, and the system's observable properties are described
		by self-adjoint operators acting on that space. %All vectors represent possible   system system states and all self-adjoint operators represent possible system   observables.
	\end{post}

	\begin{remark}
		All proofs, except one, should not depend on the dimensionality of
		the space. The exception is proposition \ref{prop_fundProj} which
		works if the basis for the space is finite since we are constructing
		the map one component at a time. It should be possible to extend it
		using weak convergence in the case of countable basis
		(i.e. separable Hilbert spaces) by creating a sequence that
		converges to each vector. We are not specifically addressing this
		case in this paper.
	\end{remark}

	\begin{post}\label{post_measurements}
		The probability that a measurement of a property $X$, described by the operator with
		spectral decomposition $X = \sum_{x,i }x \frac{| x_i \> \< x_i |}{\< x_i | x_i \>}$  where $i$ is a degeneracy index, returns a value $x$ depends only on $X$ and on the state of the system $\pj{\psi}$ and is given by $P(x|\pj{\psi})=\sum_i \frac{\<\psi| x_i \> \< x_i |\psi\>}{\< \psi | \psi \>\< x_i | x_i \>}$
		(Born rule).\end{post}

	\begin{defn}[Compatible states]\label{def_compatible}
		Let A and B be two systems. Let ${\mathcal{A}}$ and ${\mathcal{B}}$
		be their corresponding state spaces. We say two (pure) states $(\pj{a},
		\pj{b}) \in \pj{\mathcal{A}} \times \pj{\mathcal{B}}$ are compatible
		iff the respective systems can be prepared in such states at the
		same time. Formally, the proposition $\pj{a} \wedge \pj{b}$ is
		possible, which means it does not correspond to the empty set in the
		$\sigma$-algebra of the probability space\footnote{The impossible event is not an event with probability zero, rather it is an event that cannot be created at all. For example, ``the dice shows a number that is even and less than two'', or ``the electron is prepared in spin up along $x$ and also along $z$'' are impossible events.}.
	\end{defn}

	{Note that the proposition $\pj{a} \wedge \pj{b}$ refers to either
		preparation or measurement of both systems  in the respective state.
		The proposition $\pj{a} | \pj{b}$, instead, is the one that
		corresponds to preparing one system in one state and measuring the
		other system in the other state \cite{cox}. Incompatible states
		refer to non-commuting observables.}

	\begin{defn}[Preparation independence]\label{def_indep}
		Two systems are said independent iff the preparation of one does not affect the preparation of the other. Formally, all (pure) state pairs $(\pj{a}, \pj{b}) \in \pj{\mathcal{A}}\times \pj{\mathcal{B}}$ are compatible.
	\end{defn}

	\begin{prop}\label{prop_singleBorn}
		Given two systems, each prepared independently in their own state, the probability of measuring a value for one system depends only on the preparation of that system. That is, $P(\pj{a_1}|\pj{a_2}\wedge \pj{b})=P(\pj{a_1}|\pj{a_2})$.
	\end{prop}
	\begin{proof}
		We first note that, by postulate \ref{post_measurements}, the probability of measuring a value for one system depends only on the preparation of that system, which means that it is independent of the properties of any other system. Therefore $P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = \frac{\<a_1| a_2 \> \< a_2 |a_1\>}{\< a_1 | a_1 \>\< a_2 | a_2 \>} = P(\pj{a_1} | \pj{a_2})$
	\end{proof}

	\begin{defn}[Composite systems]\label{def_comp}
		Let A and B be two systems. The composite system C of A and B is formed by the simple collection of those and only those two systems, in the sense that it satisfies the following two requirements.
		\begin{enumerate}
			\item Every preparation of both subsystems is a preparation of the
			composite. Formally, let $\pj{\mathcal{C}}$ be the state space
			for C, there exists a map (not yet specified)
			$M:\pj{\mathcal{A}}\times\pj{\mathcal{B}}\to\pj{\mathcal{C}}$
			such that, for any compatible pair of (pure) states $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$, the proposition $\pj{a} \wedge \pj{b}$ is equivalent to the (pure) state $M(\pj{a},\pj{b}) \in \pj{\mathcal{C}}$ where $M$ returns the state of the composite system where the subsystems were prepared in the given states. In other words, $\pj{a} \wedge \pj{b}$ and $M(\pj{a},\pj{b})$ correspond to the same event in probability space\footnote{We will end up proving that the map $M$ leads to the tensor product.}.
			\item Every preparation of the composite gives a non-trivial measurement on the components. Formally, for every $\pj{c} \in \pj{\mathcal{C}}$, we can find at least $\pj{a} \in \pj{\mathcal{A}}$ and $\pj{b} \in \pj{\mathcal{B}}$ such that $P(\pj{a} \wedge \pj{b}|\pj{c})\neq 0$. 
		\end{enumerate}
		Requirement 1 ensures that the composite system contains all the properties of the components. Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components.
	\end{defn}

	\begin{prop}[Span surjectivity, H3]\label{prop_spanSurj}
		The map $M :
		\pj{\mathcal{A}}\times\pj{\mathcal{B}} \to \pj{\mathcal{C}}$ is span
		surjective, meaning that the span of the image coincides with the
		whole space. That is $Sp(\{ c \in \mathcal{C} \, | \, \pj{c} \in
		M(\pj{\mathcal{A}}, \pj{\mathcal{B}})\}) = \mathcal{C}$.
	\end{prop}
	\begin{proof}
		Consider $I=\{ c \in \mathcal{C} \, | \, \pj{c} \in M(\pj{\mathcal{A}}, \pj{\mathcal{B}})\}$ and its span. This forms a subspace of $\mathcal{C}$. By requirement 2 of \ref{def_comp}, for any $c \in \mathcal{C}$ we can always find $a \in \mathcal{A}$ and $b \in \mathcal{B}$ such that $P(\pj{a} \wedge \pj{b} | \pj{c} )\neq 0$. This means there is no element in $\mathcal{C}$ that is orthogonal to $Sp(I)$, therefore $Sp(I)$ must cover the whole $\mathcal{C}$.
	\end{proof}

	\begin{prop}[Totality, H1]\label{prop_totality}
		The map $M$ is in general a partial function.\footnote{A partial
			function is one that is not defined on the full domain. For
			example, $\sqrt(x)$ is a partial function since is not defined for
			$x<0$.} However, if A and B are independent, $M$ is a total function.\footnote{A total function is one that is defined on the full domain. For example, $x^2$ is a total function since it is defined for any $x$.}
	\end{prop}
	\begin{proof}
		As $M(\pj{a},\pj{b})$ is defined only if $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$ are a compatible pair of pure states, it is not defined on pairs that are not compatible. If the two systems are independent, however, all pairs are allowed and $M$ is a total function.
	\end{proof}

	\begin{remark}
		As noted in \ref{def_comp}, if $\pj{a}$ and $\pj{b}$ are incompatible, $\pj{a} \wedge \pj{b}=\emptyset$ corresponds to the impossible event (i.e.~the empty set in the $\sigma$-algebra). This is not a state, and therefore $M(\pj{a},\pj{b})$ is not defined on incompatible pairs.

		However, in the end we will construct a map $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ on the vector spaces. There the zero vector plays the role of the impossible event. Therefore independent systems will map each pair to a non-zero element of the tensor product, while systems that are not independent will map incompatible states to the zero vector (e.g. the composite state of two electrons will exclude the cases where both electrons are in the same state).
	\end{remark}

	\begin{prop}[Statistical independence]\label{prop_statInd}
		Let $\pj{\mathcal{A}}$ and $\pj{\mathcal{B}}$ be the state spaces of two quantum systems and $\pj{\mathcal{C}}$ be the state space of their composite. The map $M : \pj{\mathcal{A}} \times \pj{\mathcal{B}} \to \pj{\mathcal{C}}$ is such that:
		\begin{align}
		P(M(\pj{a_1},\pj{b}) | M(\pj{a_2},\pj{b})) = P(\pj{a_1} | \pj{a_2})  \\ P(M(\pj{a},\pj{b_1}) | M(\pj{a},\pj{b_2})) = P(\pj{b_1} | \pj{b_2})
		\end{align}
		for all $a, a_1, a_2 \in \mathcal{A}$ and $b, b_1, b_2 \in \mathcal{B}$
	\end{prop}
	\begin{proof}
		By \ref{prop_singleBorn} we have $P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = P(\pj{a_1} | \pj{a_2})$ and similarly $P(\pj{b_1} | \pj{a} \wedge \pj{b_2}) = P(\pj{b_1} | \pj{b_2})$.  Using standard probability rules and remembering that $M(\pj{a},\pj{b})\equiv \pj{a} \wedge \pj{b}$ by \ref{def_comp}, we have $P(M(\pj{a_1},\pj{b}) | M(\pj{a_2},\pj{b})) = P(\pj{a_1} \wedge \pj{b} | \pj{a_2} \wedge \pj{b})  = P(\pj{b} | \pj{a_2} \wedge \pj{b}) P(\pj{a_1} | \pj{a_2} \wedge \pj{b} \wedge \pj{b})  = P(\pj{b} | \pj{b}) P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = P(\pj{a_1} | \pj{a_2})$, since trivially $P(\pj{b}|\pj{b})=1$. Similarly $P(M(\pj{a},\pj{b_1}) | M(\pj{a},\pj{b_2})) =P(\pj{a} \wedge \pj{b_1} | \pj{a} \wedge \pj{b_2}) = P(\pj{b_1} | \pj{b_2})$
	\end{proof}

	\begin{prop}[Fundamental theorem of projective geometry]\label{prop_fundProj}
		Let $X$ and $Y$ be two complex Hilbert spaces and $\pj{X}$ and $\pj{Y}$ their respective projective spaces. Let $M : \pj{X} \to \pj{Y}$ be a map such that $P(\pj{v}|\pj{w}) = P(M(\pj{v})|M(\pj{w}))$. Then we can find, up to a total phase, a unique map $m : X \to Y$ such that $\pj{m(v)}=M(\pj{v})$. Moreover, $m$ is either linear, $\<v|w\> = \<m(v)|m(w)\>$, or anti-linear, $\<v|w\> = \<m(w)|m(v)\>$.
	\end{prop}

	\begin{remark}
		The above proposition is, for the most part, an adaptation of the fundamental theorem of projective geometry \cite{fun}. The conservation of the probability imposes the semi-linear map to be either linear or anti-linear (i.e. conjugate-linear).
	\end{remark}

	\begin{proof}
		First we note that, given an orthonormal basis $\{e_i\}_{i \in I}$ over $X$, we can use $M$ to construct a corresponding basis over $Y' \subseteq Y$ where $\pj{Y'} = M(\pj{X})$. In fact, for each $\pj{e_i}$, pick a unit $u_i \in M(\pj{e_i})$. We have $\delta_{ij} = |\<e_i | e_j \>|^2 = P(e_i | e_j) = P(\pj{e_i}|\pj{e_j}) = P(M(\pj{e_i})|M(\pj{e_j})) = P(\pj{u_i}|\pj{u_j})= |\<u_i | u_j \>|^2$. The set $\{u_i\}_{i \in I}$ spans the entire $Y'$ since for all $y \in Y'$ we can find $x \in X$ and at least one $u_i$ such that $|\<y | u_i \>|^2 = P(\pj{y}|\pj{u_i}) = P(M(\pj{x})|M(\pj{e_i})) = P(x | e_i) = |\<x | e_i \>|^2 \neq 0$. Note that we have an arbitrary choice for each $u_i$, since we have to pick a vector from the unit circle (i.e. a phase for each basis vector). This corresponds to a choice of gauge.

		We also note that the map is colinear, meaning that if $U_X, V_X \subseteq X$ are two subgroups such that $U_X \subset V_X$, then $U_Y, V_Y \subseteq Y$ such that $\pj{U_Y} = M(\pj{U_X})$ and $\pj{V_Y} = M(\pj{V_X})$ are subgroups of $Y$ and $U_Y \subset V_Y$. In fact, take a basis $\{e_i\}_{i \in I} \subset X$ such that $\{e_i\}_{i \in I_U \subset I} \subset \{e_i\}_{i \in I_V \subset I}$ are bases for $U_X$ and $V_X$ respectively. An element of $X$ belongs to $U_X$ if and only if it is not orthogonal only to elements of the basis of $U_X$ and belongs to $V_X$ only if it not orthogonal only to elements of the basis of $V_X$. As the map $M$ preserves orthogonality, these relationships are preserved by the map. Therefore $U_Y$ and $V_Y$ are subgroups of $Y$ such that $U_Y \subset V_Y$.

		We now use the gauge freedom to redefine the basis such that for all $i$ we have $M(\pj{e_i}) = \pj{v_i}$ and $M(\pj{e_1 + e_i}) = \pj{v_1 + v_i}$. Let $v_1 = u_1$. This is the only arbitrary choice we make, and corresponds to the choice of a global phase. For each $i>1$, consider $e_1 + e_i$. This will belong to the subspace $Sp(e_1, e_i)$. This subspace, when mapped through $M$, will give us the subspace spanned by $v_1$ and $u_i$. That is, $M(\pj{Sp(e_1, e_i)})=\pj{Sp(v_1, u_i)}$. This means we can find a unique $k \in \mathbb{C}$ such that $M(\pj{e_1+e_i})=\pj{v_1+ku_i}$. We fix $v_i = k u_i$. Note that $P(\pj{e_1} | \pj{e_1+e_i}) = \frac{1}{2}=  P(\pj{e_i} | \pj{e_1+e_i})=P(\pj{v_1} | \pj{v_1+k u_i})=P(\pj{u_i} | \pj{v_1+k u_i})$. Therefore $|k| = 1$ and $k u_i = v_i$ is a unit vector.

		Now we want to show that $M(\pj{e_1 + c e_i})=\pj{v_1 + \tau_i(c)
			v_i}$ where either $\tau_i(c) = c$ or $\tau_i(c) = c^\dagger$. For
		each $i$, consider $w = e_1 + c e_i \in Sp(e_1, e_i)$. Since
		$M(\pj{w}) \subset \pj{Sp(v_1, v_i)}$, there must be a $\tau_i(c)$
		such that $\pj{v_1 + \tau_i(c) v_i} = M(\pj{w})$. Since we must have
		$P(\pj{e_i} | \pj{w}) = P(\pj{v_i} | M(\pj{w}))$ and $P(\pj{e_1 +
			e_i} | \pj{w}) = P(\pj{v_1 + v_i} | M(\pj{w}))$, we must have
		$|c| = |\tau_i(c)|$ and $cos(arg(c)) = cos(arg(\tau_i(c)))$ for any
		$c$. This means that either $\tau_i(c) = c$ or $\tau_i(c) =
		c^*$.

		Next we want to show that $\tau_i(c) = \tau_j(c)$ for all pairs $(i,j)$. That is, either we have to take the complex conjugate of all components or of none. Consider $e_i - e_j$. We have $\pj{e_i - e_j} \subset \pj{Sp(e_i, e_j)}$ and, for any $c \in \mathbb{C}$, $\pj{e_i - e_j} \subset \pj{Sp(e_1 + c e_i, e_1 + c e_j)}$. By construction, we have $M(\pj{e_i - e_j})\subset \pj{Sp(v_i, v_j)}$ and $M(\pj{e_i - e_j})\subset \pj{Sp(v_1 + \tau_i(c) v_i, v_1 + \tau_j(c) v_j)}$. Therefore $M(\pj{e_i - e_j}) = \pj{Sp(v_i, v_j)} \cap \pj{Sp(v_1 + \tau_i(c) v_i, v_1 + \tau_j(c) v_j)} = \pj{\tau_i(c) v_i - \tau_j(c) v_j}$. This means that, for all $c$, $\tau_i(c) = \tau_j(c)$.

		Now we show that for all $c_2, ..., c_n \in \mathbb{C}$ we have $M(\pj{e_1 + c_2 e_2 + ... + c_n e_n}) = \pj{v_1 + \tau(c_2) v_2 + ... + \tau(c_n) v_n}$. We prove this by induction. If only the first two components are non-zero, we have  $M(\pj{e_1 + c_2 e_2}) = \pj{v_1 + \tau(c_2) v_2}$ by construction. Let $2 < p \leq n$. If we assume $M(\pj{e_1 + c_2 e_2 + ... + c_{p-1} e_{p-1}}) = \pj{v_1 + \tau(c_2) v_2 + ... + \tau(c_{p-1}) v_{p-1}}$, then $M(\pj{e_1 + c_2 e_2 + ... + c_p e_p}) \subset M(\pj{Sp(e_1 + c_2 e_2 + ... + c_{p-1} e_{p-1}, e_p})) = \pj{Sp(v_1 + c_2 v_2 + ... + c_{p-1} v_{p-1}, v_p)}$. This means that there exists $k_p \in \mathbb{C}$ such that $M(\pj{e_1 + c_2 e_2 + ... + c_p e_p}) = \pj{v_1 + c_2 v_2 + ... + c_{p-1} v_{p-1} + k_p v_p)}$. But we also have $M(\pj{e_1 + c_2 e_2 + ... + c_p e_p}) \subset M(\pj{Sp(e_1 + c_p e_p, c_2 e_2 + ... + c_{p-1} e_{p-1}})) = \pj{Sp(v_1 + \tau(c_p) v_p, c_2 v_2 + ... + c_{p-1} v_{p-1}, v_p)}$. The only way this can work is if $k_p = \tau(c_p)$.

		We also need to show the above works when there is no component on the first element of the basis. That is, for all $c_2, ..., c_n \in \mathbb{C}$ we have $M(\underbar{$c_2 e_2 + ... + c_n e_n$}) = \underbar{$\tau(c_2) v_2 + ... + \tau(c_n) v_n$}$. First note that $M(\underbar{$c_2 e_2 + ... + c_n e_n$}) \subset M(\underbar{$Sp(e_2, ..., e_n)$)} = \underbar{$Sp(v_2, ..., v_n)$}$. Also note that $M(\underbar{$c_2 e_2 + ... + c_n e_n$}) \subset M(\underbar{$Sp(e_1, e_1 + c_2 e_2 + ... + c_n e_n)$)} = \underbar{$Sp(v_1, v_1 + \tau(c_2) v_2 + ... + \tau(c_n) v_n)$}$. The only way this can work is if $M(\underbar{$c_2 e_2 + ... + c_n e_n$}) = \underbar{$\tau(c_2) v_2 + ... + \tau(c_n) v_n$}$.

		We can now define $m : X \to Y$ such that $m(e_i) = v_i$ for all $i$ and $m(\sum_{i \in I} c_i e_i) = \sum_{i \in I} \tau(c_i) v_i$. This means $\underbar{$m(\sum_{i \in I} c_i e_i)$} = M(\underbar{$\sum_{i \in I} c_i e_i$})$. Moreover, if $\tau(c) = c$ we have $\<m(\sum_{i \in I} c_i e_i)|m(\sum_{j \in I} d_j e_j)\> = \<\sum_{i \in I} c_i v_i|\sum_{j \in I} d_j v_j\> = c_i^* d_j \delta_{ij} = \<\sum_{i \in I} c_i e_i|\sum_{j \in I} d_j e_j\>$. On the other hand, if $\tau(c) = c^*$ we have $\<m(\sum_{i \in I} c_i e_i)|m(\sum_{j \in I} d_j e_j)\> = \<\sum_{i \in I} c_i^* v_i|\sum_{j \in I} d_j^* v_j\> = \<\sum_{j \in I} d_j v_j|\sum_{i \in I} c_i v_i\> = d_i^* c_j \delta_{ij} = \<\sum_{j \in I} d_j e_j|\sum_{i \in I} c_i e_i\>$.
	\end{proof}

	\begin{remark}
		The fact that the proposition identifies either a linear map or an anti-linear (i.e.~conjugate-linear)  has a clear physical interpretation. Since, for a Hilbert space, the conjugate vector space is equivalent to the dual space, a conjugate-linear map is one that preserves the inner product but maps ket vectors into bra vectors. Looking ahead, the above result does not exclude a composition map similar to the tensor product, but that maps the kets of one or both subsystems into bras in the composite system.

		Such a conjugate-linear map is physically indistinguishable from a map that maps kets into kets: the physics is only contained in the probabilities which are the square moduli of inner products. The idea is that, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear. With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.

		Note that this unnecessary subtlety could in principle be avoided by
		reformulating quantum mechanics in terms of quantum states given by
		density matrices $\rho=|\psi\>\<\psi|$ (which contain both  kets and
		bras), as is done, for example in \cite{ozawa,holevo}. In this paper
		we employed the more familiar formulation in which quantum states are
		rays in Hilbert space (identified either by kets or bras).
	\end{remark}

	\begin{prop}[Bilinearity, H2]\label{prop_bilinearity}
		Let $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ be a map such that, for all $(a,b) \in \mathcal{A} \times \mathcal{B}$, we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$. Then $m$ must be bilinear. That is:
		\begin{align}
		m(k_1a_1 + k_2a_2, b)=k_1m(a_1, b) + k_2m(a_2, b) \\
		m(a, k_1b_1 + k_2b_2)=k_1m(a, b_1) + k_2m(a, b_2)
		\end{align}
		for all $a, a_1, a_2 \in \mathcal{A}$, $b, b_1, b_2 \in \mathcal{B}$ and $k_1, k_2 \in \mathbb{C}$.
	\end{prop}

	\begin{proof}
		If we fix $b \in \mathcal{B}$, then we have $M_b : \pj{\mathcal{A}} \to \pj{\mathcal{C}}$ where $M_b(\pj{a}) = M(\pj{a}, \pj{b})$. By \ref{prop_statInd} and \ref{prop_fundProj} we can find a linear map $m_b : \mathcal{A} \to \mathcal{C}$ such that $\pj{m_b(a)} = M_b(\pj{a}) = M(\pj{a}, \pj{b})$. As this must map subspace to subspace, we must have $m(a, b) = k m_b(a)$ for some $k \in \mathbb{C}$. Since $m_b$ is linear, we have $m(k_1a_1 + k_2a_2, b)=k_1m(a_1, b) + k_2m(a_2, b)$ for any $a_1, a_2 \in \mathcal{A}$ and $k_1, k_2 \in \mathbb{C}$. We can repeat the argument fixing $a \in \mathcal{A}$, and find $m(a, k_1b_1 + k_2b_2)=k_1m(a, b_1) + k_2m(a, b_2)$ for any $b_1, b_2 \in \mathcal{B}$ and $k_1, k_2 \in \mathbb{C}$.
	\end{proof}

	\begin{prop}[Subsystems' basis gives composite system
		basis]\label{prop_basis}
		Let $\{a_i\}_{i\in I}$ and $\{b_j\}_{j \in J}$ be bases of $\mathcal{A}$ and $\mathcal{B}$ respectively, then a set of unit vectors $\{e_{ij}\}_{(i,j) \in I \times J} \subset \mathcal{C}$  such that $e_{ij} \in M(\underbar{$a_i$}, \underbar{$b_j$})$ forms a basis for $\mathcal{C}$.
	\end{prop}

	\begin{proof}

		Since $M$ is a map on the projective spaces, it maps spans to spans.
		Since the span of the basis of $\mathcal{A}$ and $\mathcal{B}$ is
		the whole space, then the span of the image of the basis is the
		whole image of $M$. By \ref{prop_spanSurj}, the image of $M$
		coincides with the whole $\mathcal{C}$. Therefore, given
		$\{a_i\}_{i\in I}$ and $\{b_j\}_{j \in J}$ bases of $\mathcal{A}$
		and $\mathcal{B}$ respectively, any set of unit vectors
		$\{e_{ij}\}_{(i,j) \in I \times J} \subset \mathcal{C}$ such that
		$e_{ij} \in M(\pj{a_i}, \pj{b_j})$ spans the whole $\mathcal{C}$.

		Now consider $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))$. If $i=k$ and $j=l$ we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))= P(M(\pj{a_i},\pj{b_j})|M(\pj{a_i},\pj{b_j})) = 1$. If $i\neq k$, we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l})) = P(\pj{a_i} \wedge \pj{b_j} | \pj{a_k} \wedge \pj{b_l}) \leq P(\pj{a_i} | \pj{a_k} \wedge \pj{b_l})$. By \ref{prop_singleBorn} we have $P(\pj{a_i} | \pj{a_k} \wedge \pj{b_l}) = P(\pj{a_i} | \pj{a_k}) = 0$ since $a_i$ and $a_k$ are different elements of an orthogonal basis. Therefore we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))=\delta_{ik}\delta_{jl}$ which means $\<e_{ij} | e_{kl} \> = \delta_{ik}\delta_{jl}$.

		The elements $e_{ij}$ form a set of orthonormal vectors that span the whole space and are therefore a basis.
	\end{proof}

	\begin{thrm}[Composite system theorem]\label{theo}
		The state space of a composite system of independent systems is given by the tensor product of the spaces of the component systems.
	\end{thrm}

	\begin{proof}
		We are looking for a map $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ such that, for all $(a,b) \in \mathcal{A} \times \mathcal{B}$ we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$. We saw in \ref{prop_bilinearity} that if $m$ exists, it must be bilinear.

		Now we show that, if $m$ exists, then $\mathcal{C} \cong \mathcal{A}
		\otimes \mathcal{B}$ (where $\cong$ indicates an isomorphism) and $m :
		\mathcal{A} \times \mathcal{B} \to \mathcal{A} \otimes \mathcal{B}$ is
		the standard map from the scalar product to the tensor product. As $m
		: \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a bilinear
		operator, by the universal property of the tensor product we can find
		a linear operator $\hat{m} : \mathcal{A} \otimes \mathcal{B} \to
		\mathcal{C}$ such that $m(a, b) = \hat{m}(a \otimes b)$. By
		\ref{prop_basis} the set $\{ m(a_i, b_j)\}_{(i,j) \in I \times J}$
		forms a basis since $ m(a_i, b_j) \in M(\pj{a_i}, \pj{b_j})$ for all
		$(i, j)$, therefore $\hat{m}(\{ a_i \otimes b_j\}_{(i,j) \in I \times
			J})$ also forms a basis since $\hat{m}(a_i\otimes b_j)=m(a_i, b_j)$.
		By \ref{prop_totality}, each $m(a_i,b_j)$ will correspond to a unit
		vector in $\mathcal{C}$. We have $\<\hat m(a_i\otimes b_j)| \hat
		m(a_k\otimes b_\ell)\>_\mathcal{C} =\<m(a_i, b_j)| m(a_k,
		b_\ell)\>_\mathcal{C} = \delta_{ik}\delta_{j\ell} = \<a_i\otimes b_j|
		a_k \otimes b_\ell\>_{\otimes}$. The function $\hat{m}$, then,
		preserves the inner product across all elements of the basis and is
		therefore an isomorphism for Hilbert spaces. We have $\mathcal{C}
		\cong \mathcal{A} \otimes \mathcal{B}$ and $m(a, b) = \hat{m}(a
		\otimes b) \cong a \otimes b$.

		Given that the tensor product map exists and it satisfies all the properties $m$ must satisfy, then $m$ exists and it is the tensor product.
	\end{proof}

	To conclude, we give a schematic outline of the logical implications
	that led us to the result. This is an expanded version of Fig.~1 of
	the main paper:
	\begin{enumerate}
		\item P\ref{post_state}: states and observables postulate.

		\item P\ref{post_measurements}: Born rule (measurement postulate).

		\item Def \ref{def_indep}: Preparation independence: systems are independent if the preparation of one does not affect the other.

		\item P\ref{post_measurements} $\Rightarrow$ \ref{prop_singleBorn}: the outcome probabilities depend only on the
		inner product.

		\item Def \ref{def_comp}: Composite system definition: A composite system is a
		collection of the subsystems (i.e.~all compatible states give a preparation) and only of the subsystems
		(i.e.~all composite preparations give non-trivial measurements on the
		subsystems).

		\item P\ref{post_measurements} + Def \ref{def_comp} $\Rightarrow$ \ref{prop_spanSurj} (H3): Span
		surjectivity (all composiste $\cal C$ are superpositions of ${\cal A}$ and ${\cal B}$).

		\item Def \ref{def_indep} + Def \ref{def_comp} $\Rightarrow$ \ref{prop_totality} (H1):
		Totality (all possible state pairs of the subsystems correspond to a state of the composite).

		\item P\ref{post_state} + \ref{prop_singleBorn} +  Def \ref{def_comp} $\Rightarrow$ \ref{prop_statInd}: Statistical
		independence (if one subsystem does not change, the probability on the composite system is given by the probability of the subsystem that changes).

		\item \ref{prop_fundProj}: Fundamental theorem of projective geometry (preserving square of inner product leads to unique linear map)

		\item \ref{prop_statInd} + \ref{prop_fundProj} $\Rightarrow$ \ref{prop_bilinearity} (H2) composition map on vector spaces is bilinear.

		\item P\ref{post_measurements} + Def \ref{def_comp} + \ref{prop_spanSurj} (H3) $\Rightarrow$ \ref{prop_basis}: Basis
		carries over from subsystems to composite

		\item \ref{prop_totality} + \ref{prop_bilinearity} + \ref{prop_basis} $\Rightarrow$ \ref{theo}: the composition map is the tensor product.
	\end{enumerate}
\end{document}

### 2020-03-31 Articles/2019-QuantumComposition/Source/QuantumComposition.tex
Version put in the archive

@@
-\documentclass{article}
-\usepackage[utf8]{inputenc}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% The four postulates of quantum mechanics are three
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\documentclass[aps,prl,amsmath,amssymb,twocolumn,nofootinbib]{revtex4}
+%\documentclass[twocolumn,aps,showpacs,prl,groupedaddress]{revtex4}
+%\usepackage{amssymb,stackengine}
 \usepackage{amssymb}
-\usepackage{amsmath}
+\usepackage{color}
+\usepackage{graphicx}
+\usepackage{epsfig,amssymb,amsmath,amsthm}
+%\usepackage[active]{srcltx}
+%\usepackage[hypertex,linkcolor=red]{hyperref}
 
-\title{Quantum system composition}
-\author{gabriele.carcassi }
-\date{October 2019}
+\usepackage{color}
+\usepackage{cancel}
+\usepackage{tikz-cd}
 
+\theoremstyle{plain}
+\newtheorem{thrm}{Theorem}[section]
+\newtheorem{form}[thrm]{Formalization}
+\newtheorem{prop}[thrm]{Proposition}
+\newtheorem{lem}[thrm]{Lemma}
+
+\theoremstyle{definition}
+\newtheorem{defn}[thrm]{Definition}
+\newtheorem{post}{Postulate}[]
+\renewcommand*{\thepost}{(\alph{post})}
+
+\theoremstyle{remark}
+\newtheorem*{remark}{Remark}
+
+\newcommand{\blue}{\color{blue}}  %NON LINEAR
+\newcommand{\red}{\color{red}}
+\newcommand{\cyan}{\color{cyan}}
+\newcommand{\green}{\color{green}} %THEORY quantum state estimation
+\newcommand{\yellow}{\color{yellow}} %THEORY quantum channel estimation
+
+\DeclareMathOperator{\spn}{span}
+
+\newcommand{\pj}[1] {\underbar{$#1$}}
+%\newcommand{\pj}[1] {\overline{#1}}
+
+\def\>{\rangle}
+\def\<{\langle}
+\def\ca{_{\cal A}}
+\def\cb{_{\cal B}}
+\def\cc{_{\cal C}}
+%\def\comment#1{}
+\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
+\def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
+\def\labell#1{\label{#1}}
+%\def\labell#1{\label{#1}{\mbox{{\tiny #1}}}}
+%\def\section#1{{\par\em #1:--- }}
+\def\sh{\mbox{sh}}
+\def\iden{\openone}
 \begin{document}
 
+\title{The four postulates of quantum mechanics are three}
+\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^2$ and Christine A. Aidala$^1$
+}\affiliation{\vbox{1.~Physics Department, University of Michigan, 450 Church Street,
+	Ann Arbor, MI 48109-1040,
+	United States}\\
+  \vbox{2.~Dip.~Fisica and INFN Sez.\ Pavia, University
+    of Pavia, via Bassi 6, I-27100 Pavia, Italy}}
+\begin{abstract}
+  The tensor product postulate of quantum mechanics states that the
+  Hilbert space of a composite system is the tensor product of the
+  components' Hilbert spaces. All current formalizations of quantum
+  mechanics that do not contain this postulate contain some equivalent
+  postulate or assumption (sometimes hidden). Here we give a natural
+  definition of composite system as a set containing the component
+  systems and show how one can logically derive the tensor product
+  rule from the state postulate and from the measurement postulate. In
+  other words, our paper reduces by one the number of postulates
+  necessary to quantum mechanics.
+\end{abstract}
+\pacs{}
+% Measurement theory (quantum mechanics), 03.65.Ta Mechanics quantum,
+% 03.65.-w noise quantum, 42.50.Lc quantum information, 03.67.Ac
+% Quantum information, 03.67.-a Quantum fluctuations, 42.50.Lc quantum
+% mechanics, 03.65.Ta quantum optics, 42.50.Gy
 \maketitle
 
-\section{First try}
+The tensor product postulate does not appear in all axiomatizations of
+quantum mechanics: it has even been called ``postulate 0'' in some
+literature \cite{zurek}. A widespread belief is that it is a direct
+consequence of the superposition principle, and it is hence not a
+necessary axiom. {\em This belief is mistaken}: the superposition
+principle is encoded into the quantum axioms by requiring that the
+state space is a {\em linear} vector space. This is, by itself,
+insufficient to single out the tensor product, as other linear
+products of linear spaces exist, such as the direct product, the
+topological product or the direct sum of vector spaces, which are used
+in classical mechanics to combine state spaces of linear systems.
+This belief may have arisen from the seminal book of Dirac
+\cite{diracbook}, who introduces tensor products (Chap.~20) by
+appealing to linearity. However, he adds the seemingly innocuous
+request that the product among spaces be distributive (rather,
+bilinear), which is equivalent to postulating tensor products (or
+linear functions of them). This is not an innocuous request. For
+example it does not hold where the composite vector space of two
+linear spaces is described by the direct product, e.g.~in classical
+mechanics, for two strings of a guitar: it is not distributive.
+[General classical systems, not only linear ones, are also composed
+through the direct product.] Of course, Dirac is not constructing an axiomatic
+formulation, so his `sleight of hand' can be forgiven. In contrast, von
+Neumann (\cite{vonneumannbook} Chap.~VI.2, also \cite{jauch})
+introduces tensor products by noticing that this is a natural choice
+in the position representation of wave mechanics (where they were
+introduced in \cite{weyl,epr}), and then {\em explicitly postulates}
+them in general: ``This rule of transformation is correct in any case
+for the coordinate and momentum operators [...] and it conforms with
+the [observable axiom and its linearity principles], we therefore
+postulate them generally.''  \cite{vonneumannbook}.  More mathematical
+or conceptually-oriented modern formulations
+(e.g.~\cite{ozawa,masanes,nielsenchuang}) introduce this postulate explicitly.
+An interesting alternative is provided in
+\cite{ballentinebook,ballentinepaper}: after introducing tensor
+products, Ballentine verifies a posteriori that they give the correct
+laws of composition of probabilities. Similarly, Peres uses
+relativistic locality \cite{peres}. While these procedures seemingly
+bypass the need to postulate the tensor product, they do not guarantee
+that this is the {\em only} possible way of introducing composite
+systems in quantum mechanics. In the framework of quantum logic,
+tensor products arise from some additional conditions \cite{matolcsi}
+which (in contrast to what is done here) are not connected to the
+other postulates. A similar approach was followed in \cite{aerts}
+where tensor products were obtained by specifying some additional
+physical requirements. In quantum field theory one tends to avoid
+problems connected with tensor products of infinite dimensional spaces
+by focusing on algebraic commutation structures, e.g.~\cite{giddins}.
+In particular, the recent MIP*=RE result \cite{mipre} implies that, in
+infinite dimensions, the tensor product is strictly less
+computationally powerful than the commutation structures, emphasizing
+the difference among these two structures, at least for the
+infinite-dimensional case. We will consider the non-relativistic
+setting here.
 
-Postulate.
-The state of the system is represented by a unit vector $a$ in a vector space $\mathcal{A}$ with inner product $< \cdot, \cdot >_\mathcal{A} : \mathcal{A} \times \mathcal{A} \to \mathbb{C}$.
+In this paper we derive the tensor product postulate (which, hence,
+loses its status of postulate) from two other postulates of quantum
+mechanics: the state postulate and the measurement postulate. We start
+from the natural definition of a composite system as the set of two
+(or more) quantum systems, in the sense that the composite system is
+made of system $A$ {\em and} (joined with) system $B$, namely a set
+whose elements are the two systems and nothing else. We will focus on
+kinematically-independent systems, namely no superselection rules or
+other restrictions to the state space are present: it is possible to
+prepare each subsystem of a composite system in a state that is
+independent of the other systems (preparation independence).  This is
+the only case in which the tensor product can be properly employed
+\cite{susskind,zanardi,zanardilloyd}: the Hilbert space of composite
+systems that have restrictions is {\em not} the tensor product of the
+component spaces, but a subspace of it (e.g.~the anti-symmetric
+subspace for fermions). Typically this is ignored in the literature,
+since the tensor product formalism is very convenient and is often
+used also in these cases. Moreover, in accordance with the measurement
+postulate, we require that the probability distribution of the
+measurement outcomes of one system is independent of the other systems' (statistical
+independence). As detailed below, statistical independence is not an
+additional requirement: it is already contained in the measurement
+postulate.
 
-%Postulate 2.
-For each measurable quantity $q$ of the system there exists an Hermitian operator $O_q : \mathcal{A} \to \mathcal{A}$ such that the expectation value is given by $\bar{q}=<v, O_qv>_\mathcal{A}$.
+From the above definition of composite system, it follows that there
+must be a map $M$ that connects the states of the subsystems to the
+states of the composite system. A quantum state is a ray in Hilbert
+space, namely a set of vectors. The map $M$ on states 
+corresponds to a map $m$ on vectors. We then show that preparation
+independence and statistical independence imply three conditions on
+the map $m$: (H1)~totality: the map is defined on all states of the
+subsystems; (H2)~bilinearity: the map is bilinear thanks to the
+fundamental theorem of projective geometry; (H3)~span surjectivity:
+the span of the image of map coincides with the full composite Hilbert space.
+We then prove that, if these three conditions H1, H2 and H3 hold, then
+the map $m$ is the tensor product, namely the Hilbert space of the
+composite system is a tensor product of the components: the tensor
+product ``postulate'', which hence loses its status of a postulate. An
+overview of all these logical implications is given in
+Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
+argument. Refer to the Appendix for a more rigorous proof of the \emph{same} argument.
 
-Definition 1. A composite system $\mathcal{C}$ of two subsystems $\mathcal{A}$ and $\mathcal{B}$ is such that:
+%%BoundingBox: 45 418 730 588
+%(%old BoundingBox: 45 382 649 547)
+\begin{figure}[ht]
+%\epsfxsize=1.\hsize\leavevmode\epsffile{fig.eps}
+\includegraphics[width=\linewidth, trim={0.2in 5.8in 2.5in 0.1in}, clip=true]{fig.eps}
+\caption{Schematic depiction of the logical implications used
+in this paper. FTPG stands for ``Fundamental Theorem of Projective Geometry''.  \label{f:fig}}\end{figure}
+
+We use the axiomatization of quantum mechanics based on the following
+postulates (e.g.~\cite{ozawa,masanes,nielsenchuang}): (a)~The state of a
+system is described by a ray $\pj{\psi}$ corresponding to a set of
+non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
+system's observable properties are described by self-adjoint operators
+acting on that space; (b)~The probability that a measurement of a
+property $X$, described by the operator with spectral decomposition
+$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
+$x$ given that the system is in state $\pj{\psi}$ is
+$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2$ (Born rule). (c)~The state
+space of a composite system is given by the tensor product of the
+spaces of the component systems; (d)~The time evolution of an isolated
+system is described by a unitary operator acting on a vector
+representing the system state, $|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or,
+equivalently, by the Schr\"odinger equation. The rest of quantum
+theory can be derived from these axioms. While some axiomatizations
+introduce further postulates, we will be using only (a) and (b) to
+derive (c), so the above are sufficient for our aims.
+
+  As stated
+above, we will limit ourselves to kinematically-independent systems,
+where all state vectors $|\psi\>$ in the system's Hilbert space $\cal H$
+describe a valid state, {\em unconditioned on anything else}. In
+particular, we note that restrictions due to superselection rules
+arise either from practical (not fundamental) limitations on the
+actions of the experimenter \cite{susskind,zanardi,zanardilloyd} or
+from the use of ill-defined quantum systems. For example, in
+situations where indistinguishability plays a role (e.g.~in QFT), one
+cannot consider an electron as a quantum system: in this case the
+quantum system is the field. The electron is an excitation of the
+field, {\em not} a (sub)system.
+
+This implies that the ``system'' is kinematically independent from
+anything else.  We call this condition ``preparation independence''.
+[We emphasize that the kinematic independence is inequivalent to
+dynamical independence (or isolation).  Indeed if two systems
+interact, their interaction may lead to dynamical restrictions in the
+state spaces. We will not consider dynamical evolution in this paper,
+which is contained in postulate (d).]
+
+The definition of a composite system as containing {\em only} the
+collection of the subsystems means that any preparation of both
+subsystems independently must correspond to the preparation of the
+composite system. Since states are defined by postulate (a) as rays in
+the respective Hilbert spaces, there must exist a map $M : \pj{\cal A}
+\times \pj{\cal B } \to \pj{\cal C}$ that takes a pair of states for
+the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$ represent the
+projective spaces and the Cartesian product is the set of all possible
+pairs) and returns a state in the projective space $\pj{\cal C}$ for
+the composite. The map $M$ that acts on rays corresponds to a map $m:{\cal A}\times{\cal B}\to{\cal C}$ that
+acts on vectors in the Hilbert spaces $\cal A$, $\cal B$ and $\cal C$.
+Namely, $\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign
+indicates the elements in the projective space. We will prove that the
+map $m$ is the tensor product.
+
+The map $M$ must be injective: as said above, different states of the
+subsystems must correspond, by definition of composite system, to
+different states of the composite. Moreover, preparation independence
+implies that $M$, and hence $m$, must be total maps: each subsystem of
+the composite system can be independently prepared and gives rise to a
+state of the composite (condition H1).  H1 is not sufficient
+to identify the tensor product: by itself it does not even guarantee
+that the map $m$ is linear.
+
+Postulate (b) contains the connection between quantum mechanics
+and probability theory. It must then implicitly contain the
+axiomatization of probability, e.g.~see
+\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
+probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
+the joint probability of events $a$ and $b$ given $z$ is
+$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Then the events $a$ and $b$ are
+independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$ (this
+is the {\em definition} of independent events). Since the Born
+probability formula in postulate (b) contains only quantities of the
+system (the state and the observable's eigenstates {\em of the
+  system}), it implies the ``statistical independence'' between
+different systems if they are prepared independently. Namely, the
+outcome probability of one is independent of any properties of any
+other in the absence of any dynamical coupling and supposing
+independent preparation. [Of course, the dynamics may couple the state
+of the system to other systems so that the outcomes may depend on what
+happens to other systems.  Analogously, one may prepare the system in
+a way that depends on other systems. But we consider independent
+preparations and do not consider dynamics, as it is sufficient for our
+aims.] We can then formalize ``statistical independence'' of
+independently prepared systems by saying that their Born rule
+satisfies:
+\begin{eqnarray}
+  p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
+    B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
+    A})\:p(b|\psi_{\cal B})
+\labell{bindip}\;,
+\end{eqnarray}
+where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
+preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
+$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
+B$ respectively. The first equality in \eqref{bindip} embodies the
+definition of statistical independence, the second follows from the
+form of the Born rule for independently prepared systems (the
+probability depends only on the state and observables of each system
+on its own).
+
+A particular case is when one system is
+prepared in a fixed eigenstate $|b\>$ of the observable measured on it,
+$p(a\wedge b|\psi\wedge b)=p(a|\psi)\:p(b|b)=p(a|\psi)$, where the last equality follows from the Born rule since $\<b|b\>=1$. 
+  If we define $M_b(\pj{a}) = M(\pj{a},\pj{b})$, substituting the values of the
+probabilities from the Born rule, we have:
+\begin{align} &\Big|\Big\<M\left(\pj{a},\pj{b}\right)\Big|M\left(\pj{\psi},\pj{b}\right)\Big\>_{\cal C}\Big|^2
+  =\Big|\Big\<M_b\left(\pj{a}\right)\Big|M_b\left(\pj{\psi}\right)\Big\>_{\cal C}\Big|^2
+\nonumber \\&
+=\left|\<a|\psi\>_{\cal A}\right|^2
+\labell{questa},
+\end{align}
+where the first and second terms contain the inner product in the composite
+space $\cal C$. [This is not a new assumption: it follows from the
+measurement postulate (b) for the composite system.] This means that,
+when one subsystem is prepared in an eigenstate of what is measured
+there, the state space of the other is mapped preserving the square of
+the inner product.
+
+Since the square of the inner product is preserved, orthogonality and
+the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
+colinear transformation by definition. In this case, the fundamental
+theorem of projective geometry \cite{fun} applies, which tells us that
+a unique semi-linear map $m_b$ that acts on the vectors exists in accordance with $M_b$.
+Moreover, conservation of probability further constrains it to be
+either linear or antilinear. This tells us that the corresponding $m$
+is either linear or antilinear in the first argument. Namely, if
+\eqref{questa} holds, then
+\begin{align}
+\<a|\psi\>=\<m(a,b)|m(\psi,b)\>\labell{h2}\;
+ \\\mbox{ or }
+\<a|\psi\>=\<m(\psi,b)|m(a,b)\> \labell{h2b}.
+\end{align}
+We can ignore the antilinear case \eqref{h2b}, which simply
+corresponds to the map $m$ mapping kets of one or both subsystems into
+bras before outputting the composite system (all these maps are
+trivially physically equivalent, since the dual-space description of a
+quantum system through bras is equivalent to the description using
+kets). We can now repeat the same analysis for the second argument of
+$m$ to conclude that it is a bilinear map, condition (H2).
+
+The last condition (H3) follows directly from the definition of a
+composite system. Since it is composed {\em only} of the component
+systems, for any state $c$ of the composite system, we must find at least one pair $(a, b)$ such that $p(a\wedge b | c)\neq 0$. It follows that the map $m$ is span-surjective: namely the
+span of the map applied to all states in the component systems spans
+the composite system state space. In other words, the composite does
+not contain states that are totally independent of (i.e.~orthogonal
+to) the states of the components.
+
+We have obtained the conditions H1, H2 and H3 from the state postulate
+(a), the measurement postulate (b) and the definitions of composite and independent
+systems. We now prove that these three conditions imply that the (up
+to now unspecified) composition rule $m$ is the tensor product. More
+precisely, given a total, span-surjective, bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ that preserves the square of the inner
+product, we find that $\cal C $ is equivalent to
+$\cal A \otimes \cal B $ and that $m=\otimes$.
+
+Proof. Step 1: the bases of the component systems are mapped to a
+basis of the composite system. Because of totality property (H1) and
+because the square of the inner product is preserved, we can conclude
+that, given two orthonormal bases $\{|a_i\>\}\in{\cal A}$ and
+$\{|b_j\>\}\in{\cal B}$,
+$|\<m(a_i,b_j)|m(a_k,b_\ell)\>|^2=\delta_{ik}\delta_{j\ell}$, namely
+$\{|m(a_i,b_j)\>\}$ is an orthonormal set in $\cal C$.  Moreover, the
+surjectivity property (H3) guarantees that in $\cal C$ no vectors are
+orthogonal to this set. This implies that it is a basis for $\cal C$.
+
+Step 2: use the universal property. The tensor product is uniquely
+characterized, up to isomorphism, by a universal property regarding
+bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
+tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
+map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the property
+than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ factors
+through $T$ uniquely.  This means that there exists a {\em unique}
+$\hat m$, dependent on $m$, such that $\hat m \circ T=m$.  In other
+words, the following diagram commutes:
+\begin{center}
+\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{m}"] \\
+& \mathcal{C}
+\end{tikzcd}
+\end{center}
+Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is
+a bilinear operator (property H2), thanks to the universal property of
+the tensor product we can find a unique linear operator $\hat{m} :
+\mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that $m(a, b) =
+\hat m(a \otimes b)$. The set $\{ \hat m(a_i\otimes b_j)$ with
+$|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and ${\cal B}\}$
+forms a basis for $\cal C$, since $\hat m(a_i\otimes b_j)=m(a_i,b_j)$
+and we have shown above that the latter is a basis.  Thus,
+\begin{align} 
+  &\<\hat m(a_i\otimes b_j)|\hat m(a_k\otimes b_\ell)
+\>_{\mathcal{C}}=\<m(a_i,
+  b_j)|m(a_k,b_\ell)\>_\mathcal{C} \nonumber\\& =
+  \delta_{ik}\delta_{j\ell}
+ = \<a_i\otimes
+  b_j| a_k \otimes b_\ell\>_{\otimes},
+\labell{ecco}\; 
+\end{align}
+where we used the orthonormality of the bases and the fact that
+$|a_i\otimes b_j\>$ is a basis of the tensor product space ${\cal
+  A}\otimes{\cal B}$. The function $\hat{m}$, then, is an isomorphism
+over all elements of the basis $|a_i\otimes b_j\>$. It is then an
+isomorphism between $\cal C$ and ${\cal A}\otimes{\cal B}$. Namely,
+$|m(a_i,b_j)\>_{\cal C} = |\hat{m}(a_i\otimes b_j)\>_{\cal C} \cong
+|a_i\otimes b_j\>_\otimes$, where the first two vectors (in $\cal C$)
+are isomorphic to the last vector (in the tensor product space ${\cal
+  A}\otimes{\cal B}$). Namely, $\hat{m}$ maps bijectively to the
+tensor product, which implies the isomorphism ${\cal C}\cong{\cal
+  A}\otimes{\cal B}$, i.e.~their equivalence through the map
+$\hat{m}$. Moreover, since ${\cal C}\cong{\cal A}\otimes{\cal B}$,
+then the map $m : \cal A \times \cal B \to \cal C$ is equivalent to
+the map $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$,
+i.e.~the maps $m$ and $\otimes$ are also equivalent through $\hat
+m$.$\square$
+
+\vskip 1\baselineskip A few comments on the proof: it is based on the
+universal property of the tensor product, which uniquely characterizes
+it. In Step 1 we showed that the bilinear map $m$ maps subsystems'
+bases into the composite system basis. We also know that there exists a
+tensor product map $ {T}=\otimes$ that can compose the vectors in $\cal A$
+and $\cal B$. In Step 2 we use the universal property: since $m$ is a
+bilinear map, we are assured that there exists a unique $\hat m$ such
+that $\hat m \circ T=m$. Since we show that $\hat m$ is an
+isomorphism, then $\hat{m}$ bijectively maps vectors in $\cal C$
+onto vectors in the tensor product space. Namely $m={T}=\otimes$.
+
+We conclude with some general comments. The tensor product structure
+of quantum systems is not absolute, but depends on the observables
+that are accessible \cite{zanardi,zanardilloyd}. This is due to the
+fact that an agent that has access to a set of observables will define
+quantum systems differently from an agent that has access to a
+different set of observables. Where one agent sees a single system, an
+agent that has access to less refined observables (and is then limited
+by some superselection rules) can consider the same system as composed
+of multiple subsystems. A typical example \cite{tellerbook} comes from
+quantum field theory. It is customary in basically all quantum optics
+literature to treat different modes of the radiation field (e.g.~the
+output of two lasers) as independent systems composed through the
+tensor product.  Clearly the electromagnetic field is a single system
+and an agent who is able to access an optical mode that is a linear
+combination of the two will give a quantum description for it that
+cannot easily accommodate tensor products. Similarly, an agent can
+consider two electrons as two systems, joined with the tensor product,
+whenever they are distinguishable for all practical purposes (e.g.~the
+electrons are in widely separated physical locations). Yet, in
+principle, electrons are just excitations of a field, and the `true'
+quantum system is the field, not the single electrons
+\cite{teller,tellerbook}.  So, in quantum field theory, the quantum
+systems that should be joined through tensor products are the
+different fields and {\em not} the particles, which are just
+excitations (states) of the fields. In the words of Teller
+(\cite{tellerbook}, pg.22), tensor products can be safely used only if
+there is a ``primitive thisness'', which is captured in the definition
+of system.
+
+It has been pointed out before that the quantum postulates are
+redundant: in \cite{masanes} it was shown that the measurement
+postulate (b) can be derived from the others (a), (c), (d). Here
+instead we have shown how the tensor product postulate can be
+logically derived from the state postulate (a), the measurement
+postulate (b) and a reasonable definition of independent systems, and
+we have described the logical relations among them.  Of course, we do
+not claim that this is the {\em only} way to obtain the tensor product
+postulate from the others.
+
+L.M. acknowledges useful discussions with M.~Ozawa, P.~Zanardi,
+S.~Lloyd, D.~Zeh, G.~Auletta, A.~Aldeni and funding from
+the Attract project through the Eu Horizon 2020 research and
+innovation programme under grant agreement No 777222. G.C. and C.A.A. acknowledge funding from the MCubed program of the University of Michigan.
+
+\begin{references}
+\bibitem{zurek} W.H. Zurek, Quantum Darwinism, Nature Phys. {\bf 5},
+  181 (2009).
+\bibitem{diracbook}P.A.M. Dirac, The principles of quantum mechanics,
+  (Clarendon Press, Oxford, 1966).
+\bibitem{vonneumannbook}J. von Neumann, Mathematical Foundations of
+  Quantum Mechanics (Princeton Univ.  Press, 1955).
+\bibitem{jauch}J.M. Jauch, Foundations of quantum mechanics
+  (Addison-Welsey, 1968), pg.~176.
+\bibitem{weyl} H. Weyl, Gruppentheorie und Quantenmechanik (Hirzel,
+  Leipzig, 1928); translated by H. P. Robertson, The Theory of Groups
+  and Quantum Mechanics (Methuen, London, 1931); reprinted by Dover,
+  p. 91.
+\bibitem{epr}A. Einstein, B. Podolsky, N. Rosen, Can
+  quantum-mechanical description of physical reality be considered
+  complete?, Phys. Rev. {\bf 47}, 777 (1935).
+\bibitem{ozawa}M. Ozawa, {Uncertainty relations for noise and
+    disturbance in generalized quantum measurements}, Ann. Phys.  {\bf
+    311}, 350 (2004).
+\bibitem{masanes}L. Masanes, T.D. Galley, M.P. M\" uller, The
+  measurement postulates of quantum mechanics are operationally
+  redundant, Nat. Commun. {\bf 10}, 1361 (2019).
+\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum Computation
+  and Quantum Information (Cambridge University Press, Cambridge,
+  2000).
+\bibitem{ballentinebook}L.E. Ballentine, Quantum Mechanics, a modern
+  development (World Scientific, 2014).
+\bibitem{ballentinepaper}L.E. Ballentine, Probability theory in
+  quantum mechanics, Am. J. Phys. {\bf 54}, 883 (1986).
+\bibitem{peres}A.~Peres, Classical interventions in quantum systems.
+  II. Relativistic invariance, Phys. Rev. A {\bf 61}, 022117 (2000).
+\bibitem{matolcsi} T. Matolcsi, Tensor product of Hilbert lattices and
+  free orthodistributive product of orthomodular lattices, Acta Sci.
+  Math. (Szeged), {\bf 37}, 263 (1975).
+\bibitem{aerts} D. Aerts, I. Daubechies, Physical justification for
+  using the tensor product to describe two quantum systems as one
+  joint system, Helv. Phys. Acta {\bf 51}, 661 (1979).
+\bibitem{giddins}S.B.~Giddings, Hilbert space structure in quantum
+  gravity: an algebraic perspective. J. High Energ. Phys. 2015, 1
+  (2015).% https://doi.org/10.1007/JHEP12(2015)099
+\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
+\bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
+  rule, Phys. Rev. {\bf 155}, 1428 (1967).
+\bibitem{zanardi}P. Zanardi, Virtual Quantum Subsystems, Phys. Rev.
+  Lett {\bf 87}, 077901 (2001).
+\bibitem{zanardilloyd} P. Zanardi, D.A. Lidar, S. Lloyd, Quantum
+  Tensor Product Structures are Observable Induced, Phys. Rev. Lett.
+  {\bf 92},060402 (2004).
+\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
+  press, 1961).
+\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
+\bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
+  Field Theory (Princeton Univ. Press, 1997).  
+\bibitem{teller}M. Redhead, P. Teller, Particles, Particle Labels, and
+  Quanta: The Toll of Unacknowledged Metaphysics, Found. Phys. {\bf
+    21}, 43 (1991).
+\bibitem{holevo}A. Holevo, Probabilistic and statistical aspects of
+  quantum theory, (North Holland, 1982).
+\end{references}
+
+\vskip 5\baselineskip
+
+\section{Appendix: mathematical formulation}\label{app}
+\setcounter{section}{1}
+
+The following section needs to distinguish the projective space from
+the Hilbert space itself. As this is seldom done in quantum mechanics,
+we review some concepts in that context and introduce the notation
+that we will be using. If $X$ is a Hilbert space, we denote $\pj{X}$
+the projective space. The projective space is mathematically
+constructed from the Hilbert space by removing the origin and
+quotienting by the equivalence relationship $v \sim \lambda v$, $v\in
+X$ and $\lambda\in\mathbb{C}$. A quantum state is a point in
+projective space. Each point of the projective space is called a ray,
+because for a real vector space to would correspond to a line going through
+the origin, with the origin removed. As we are in a complex space, the
+ray should be thought as a complex plane without the origin, which is
+the space of the vectors reachable from a fixed one through
+multiplication by a complex number. It can also be thought as a
+subspace of dimension one.
+
+Given a vector $v \in X$, we will denote $\pj{v}$ the ray in the
+projective space corresponding to $v$. Note that $\pj{v}$ denotes a
+quantum state, without having picked a modulus or phase. Given two or
+more vectors $v_1, ..., v_n \in X$, the subspace of $X$ they span
+(i.e.~all the vectors reached by linear combinations) is noted by
+$Sp(v_1, ..., v_n)$. Note that this subspace will correspond to a set
+of rays in the projective space, which we note as $\pj{Sp(v_1, ...,
+  v_n)}$. Given $v,w \in X$, we can write $P(v|w) = \frac{|\< v | w
+  \>|^2}{\< v | v \>\< w | w \>}$ which corresponds to the probability
+of observing $v$ given $w$ was prepared. Note that $P(v|w) = P(\lambda
+v| \mu w)$, with non-null $\lambda,\mu\in{\mathbb C}$, and therefore
+one can write $P(\pj{v} | \pj{w})\equiv P(v|w)$ as a function of the
+rays.
+
+\begin{post}\label{post_state}
+  The state of a quantum system is described by a ray $\pj{\psi} = \{
+  \alpha |\psi\> \, |$ non-null $\alpha \in
+  \mathbb{C},|\psi\>\in\mathcal{H}\}$ in a complex Hilbert space
+  $\mathcal{H}$, and the system's observable properties are described
+  by self-adjoint operators acting on that space. %All vectors represent possible   system system states and all self-adjoint operators represent possible system   observables.
+\end{post}
+
+\begin{remark}
+  All proofs, except one, should not depend on the dimensionality of
+  the space. The exception is proposition \ref{prop_fundProj} which
+  works if the basis for the space is finite since we are constructing
+  the map one component at a time. It should be possible to extend it
+  using weak convergence in the case of countable basis
+  (i.e. separable Hilbert spaces) by creating a sequence that
+  converges to each vector. We are not specifically addressing this
+  case in this paper.
+\end{remark}
+
+\begin{post}\label{post_measurements}
+The probability that a measurement of a property $X$, described by the operator with
+spectral decomposition $X = \sum_{x,i }x \frac{| x_i \> \< x_i |}{\< x_i | x_i \>}$  where $i$ is a degeneracy index, returns a value $x$ depends only on $X$ and on the state of the system $\pj{\psi}$ and is given by $P(x|\pj{\psi})=\sum_i \frac{\<\psi| x_i \> \< x_i |\psi\>}{\< \psi | \psi \>\< x_i | x_i \>}$
+(Born rule).\end{post}
+
+\begin{defn}[Compatible states]\label{def_compatible}
+  Let A and B be two systems. Let ${\mathcal{A}}$ and ${\mathcal{B}}$
+   be their corresponding state spaces. We say two (pure) states $(\pj{a},
+  \pj{b}) \in \pj{\mathcal{A}} \times \pj{\mathcal{B}}$ are compatible
+  iff the respective systems can be prepared in such states at the
+  same time. Formally, the proposition $\pj{a} \wedge \pj{b}$ is
+  possible, which means it does not correspond to the empty set in the
+  $\sigma$-algebra of the probability space\footnote{The impossible event is not an event with probability zero, rather it is an event that cannot be created at all.
+For example, ``the dice shows a number that is even and less than two'', or ``the electron is prepared in spin up along $x$ and also along $z$'' are impossible events.}.
+\end{defn}
+
+{Note that the proposition $\pj{a} \wedge \pj{b}$ refers to either
+  preparation or measurement of both systems  in the respective state.
+  The proposition $\pj{a} | \pj{b}$, instead, is the one that
+  corresponds to preparing one system in one state and measuring the
+  other system in the other state \cite{cox}. Incompatible states
+  refer to non-commuting observables.}
+
+\begin{defn}[Preparation independence]\label{def_indep}
+Two systems are said independent iff the preparation of one does not affect the preparation of the other.
+Formally, all (pure) state pairs $(\pj{a}, \pj{b}) \in \pj{\mathcal{A}}\times \pj{\mathcal{B}}$ are compatible.
+\end{defn}
+
+\begin{prop}\label{prop_singleBorn}
+  Given two systems, each prepared independently in their own state, the probability of measuring a value for one system depends only on the preparation of that system.
+That is, $P(\pj{a_1}|\pj{a_2}\wedge \pj{b})=P(\pj{a_1}|\pj{a_2})$.
+\end{prop}
+\begin{proof}
+We first note that, by postulate \ref{post_measurements}, the probability of measuring a value for one system depends only on the preparation of that system, which means that it is independent of the properties of any other system.
+Therefore $P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = \frac{\<a_1| a_2 \> \< a_2 |a_1\>}{\< a_1 | a_1 \>\< a_2 | a_2 \>} = P(\pj{a_1} | \pj{a_2})$
+\end{proof}
+
+\begin{defn}[Composite systems]\label{def_comp}
+Let A and B be two systems. The composite system C of A and B is formed by the simple collection of those and only those two systems, in the sense that it satisfies the following two requirements.
 \begin{enumerate}
-    \item given the states of the two subsystems we can construct a state in which those subsystems have that state; that is, there exists an injective map $C : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$
+    \item Every preparation of both subsystems is a preparation of the
+      composite. Formally, let $\pj{\mathcal{C}}$ be the state space
+      for C, there exists a map (not yet specified)
+      $M:\pj{\mathcal{A}}\times\pj{\mathcal{B}}\to\pj{\mathcal{C}}$
+      such that, for any compatible pair of (pure) states $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$, the proposition $\pj{a} \wedge \pj{b}$ is equivalent to the (pure) state $M(\pj{a},\pj{b}) \in \pj{\mathcal{C}}$ where $M$ returns the state of the composite system where the subsystems were prepared in the given states.
+In other words, $\pj{a} \wedge \pj{b}$ and $M(\pj{a},\pj{b})$ correspond to the same event in probability space\footnote{We will end up proving that the map $M$ leads to the tensor product.}.
+    \item Every preparation of the composite gives a non-trivial measurement on the components.
+Formally, for every $\pj{c} \in \pj{\mathcal{C}}$, we can find at least $\pj{a} \in \pj{\mathcal{A}}$ and $\pj{b} \in \pj{\mathcal{B}}$ such that $P(\pj{a} \wedge \pj{b}|\pj{c})\neq 0$.
+\end{enumerate}
+Requirement 1 ensures that the composite system contains all the properties of the components.
+Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components.
+\end{defn}
 
-    \item when the state of one of the subsystems is fixed, the state space of the composite system is equivalent to the one of the free subsystem.
-That is, given a unitary vector $b \in \mathcal{B}$ then for all $a_1, a_2 \in \mathcal{A}$:
+\begin{prop}[Span surjectivity, H3]\label{prop_spanSurj}
+  The map $M :
+  \pj{\mathcal{A}}\times\pj{\mathcal{B}} \to \pj{\mathcal{C}}$ is span
+  surjective, meaning that the span of the image coincides with the
+  whole space. That is $Sp(\{ c \in \mathcal{C} \, | \, \pj{c} \in
+  M(\pj{\mathcal{A}}, \pj{\mathcal{B}})\}) = \mathcal{C}$.
+\end{prop}
+\begin{proof}
+  Consider $I=\{ c \in \mathcal{C} \, | \, \pj{c} \in M(\pj{\mathcal{A}}, \pj{\mathcal{B}})\}$ and its span.
+This forms a subspace of $\mathcal{C}$.
+By requirement 2 of \ref{def_comp}, for any $c \in \mathcal{C}$ we can always find $a \in \mathcal{A}$ and $b \in \mathcal{B}$ such that $P(\pj{a} \wedge \pj{b} | \pj{c} )\neq 0$.
+This means there is no element in $\mathcal{C}$ that is orthogonal to $Sp(I)$, therefore $Sp(I)$ must cover the whole $\mathcal{C}$.
+\end{proof}
 
-    $$<a_1, a_2>_\mathcal{A} = <C(a_1, b), C(a_2, b)>_\mathcal{C}.$$
+\begin{prop}[Totality, H1]\label{prop_totality}
+  The map $M$ is in general a partial function.\footnote{A partial
+    function is one that is not defined on the full domain. For
+    example, $\sqrt(x)$ is a partial function since is not defined for
+    $x<0$.} However, if A and B are independent, $M$ is a total function.\footnote{A total function is one that is defined on the full domain.
+For example, $x^2$ is a total function since it is defined for any $x$.}
+\end{prop}
+\begin{proof}
+  As $M(\pj{a},\pj{b})$ is defined only if $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$ are a compatible pair of pure states, it is not defined on pairs that are not compatible.
+If the two systems are independent, however, all pairs are allowed and $M$ is a total function.
+\end{proof}
 
-    Similarly, given a unitary vector $a \in \mathcal{A}$ then for all $b_1, b_2 \in \mathcal{B}$:
+\begin{remark}
+   As noted in \ref{def_comp}, if $\pj{a}$ and $\pj{b}$ are incompatible, $\pj{a} \wedge \pj{b}=\emptyset$ corresponds to the impossible event (i.e.~the empty set in the $\sigma$-algebra).
+This is not a state, and therefore $M(\pj{a},\pj{b})$ is not defined on incompatible pairs.
 
-    $$<b_1, b_2>_\mathcal{B} = <C(a, b_1), C(a, b_1)>_\mathcal{C}.$$
+   However, in the end we will construct a map $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ on the vector spaces.
+There the zero vector plays the role of the impossible event.
+Therefore independent systems will map each pair to a non-zero element of the tensor product, while systems that are not independent will map incompatible states to the zero vector (e.g.
+the composite state of two electrons will exclude the cases where both electrons are in the same state).
+\end{remark}
 
-%\item every measurable quantity on each subsystem corresponds to a measurable quantity on the composite system with the same expectation value; that is, for every Hermitian operator $O_{q_a}^\mathcal{A} : \mathcal{A} \to \mathcal{A}$ associated with a quantity $q_a$ there exists an Hermitian operator $O_{q_a}^\mathcal{C} :\mathcal{C} \to \mathcal{C}$ such that
-%$$\bar{q_a}=<v_a, O_{q_a}^\mathcal{A}v_a>_\mathcal{A} = <C(v_a, v_b), O_{q_a}^\mathcal{C}C(v_a, v_b)>_\mathcal{C}$$
-%for all $v_a \in V_\mathcal{A}$, $v_b \in V_\mathcal{B}$.
-Similarly, for every Hermitian operator $O_{q_b}^\mathcal{B} : \mathcal{B} \to \mathcal{B}$ associated with a quantity $q_b$ there exists an Hermitian operator $O_{q_b}^\mathcal{C} :\mathcal{C} \to \mathcal{C}$ such that
-%$$\bar{q_b}=<v_b, O_{q_b}^\mathcal{B}v_b>_\mathcal{B} = <C(v_a, v_b), O_{q_a}^\mathcal{C}C(v_a, v_b)>_\mathcal{C}$$
-%for all $v_a \in V_\mathcal{A}$, $v_b \in V_\mathcal{B}$.
+\begin{prop}[Statistical independence]\label{prop_statInd}
+Let $\pj{\mathcal{A}}$ and $\pj{\mathcal{B}}$ be the state spaces of two quantum systems and $\pj{\mathcal{C}}$ be the state space of their composite.
+The map $M : \pj{\mathcal{A}} \times \pj{\mathcal{B}} \to \pj{\mathcal{C}}$ is such that:
+\begin{align}
+P(M(\pj{a_1},\pj{b}) | M(\pj{a_2},\pj{b})) = P(\pj{a_1} | \pj{a_2})  \\ P(M(\pj{a},\pj{b_1}) | M(\pj{a},\pj{b_2})) = P(\pj{b_1} | \pj{b_2})
+\end{align}
+for all $a, a_1, a_2 \in \mathcal{A}$ and $b, b_1, b_2 \in \mathcal{B}$
+\end{prop}
+\begin{proof}
+  By \ref{prop_singleBorn} we have $P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = P(\pj{a_1} | \pj{a_2})$ and similarly $P(\pj{b_1} | \pj{a} \wedge \pj{b_2}) = P(\pj{b_1} | \pj{b_2})$.
+Using standard probability rules and remembering that $M(\pj{a},\pj{b})\equiv \pj{a} \wedge \pj{b}$ by \ref{def_comp}, we have $P(M(\pj{a_1},\pj{b}) | M(\pj{a_2},\pj{b})) = P(\pj{a_1} \wedge \pj{b} | \pj{a_2} \wedge \pj{b})  = P(\pj{b} | \pj{a_2} \wedge \pj{b}) P(\pj{a_1} | \pj{a_2} \wedge \pj{b} \wedge \pj{b})  = P(\pj{b} | \pj{b}) P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = P(\pj{a_1} | \pj{a_2})$, since trivially $P(\pj{b}|\pj{b})=1$.
+Similarly $P(M(\pj{a},\pj{b_1}) | M(\pj{a},\pj{b_2})) =P(\pj{a} \wedge \pj{b_1} | \pj{a} \wedge \pj{b_2}) = P(\pj{b_1} | \pj{b_2})$
+\end{proof}
+
+\begin{prop}[Fundamental theorem of projective geometry]\label{prop_fundProj}
+    Let $X$ and $Y$ be two complex Hilbert spaces and $\pj{X}$ and $\pj{Y}$ their respective projective spaces.
+Let $M : \pj{X} \to \pj{Y}$ be a map such that $P(\pj{v}|\pj{w}) = P(M(\pj{v})|M(\pj{w}))$.
+Then we can find, up to a total phase, a unique map $m : X \to Y$ such that $\pj{m(v)}=M(\pj{v})$.
+Moreover, $m$ is either linear, $\<v|w\> = \<m(v)|m(w)\>$, or anti-linear, $\<v|w\> = \<m(w)|m(v)\>$.
+\end{prop}
+
+\begin{remark}
+  The above proposition is, for the most part, an adaptation of the fundamental theorem of projective geometry \cite{fun}.
+The conservation of the probability imposes the semi-linear map to be either linear or anti-linear (i.e.
+conjugate-linear).
+\end{remark}
+
+\begin{proof}
+  First we note that, given an orthonormal basis $\{e_i\}_{i \in I}$ over $X$, we can use $M$ to construct a corresponding basis over $Y' \subseteq Y$ where $\pj{Y'} = M(\pj{X})$.
+In fact, for each $\pj{e_i}$, pick a unit $u_i \in M(\pj{e_i})$.
+We have $\delta_{ij} = |\<e_i | e_j \>|^2 = P(e_i | e_j) = P(\pj{e_i}|\pj{e_j}) = P(M(\pj{e_i})|M(\pj{e_j})) = P(\pj{u_i}|\pj{u_j})= |\<u_i | u_j \>|^2$.
+The set $\{u_i\}_{i \in I}$ spans the entire $Y'$ since for all $y \in Y'$ we can find $x \in X$ and at least one $u_i$ such that $|\<y | u_i \>|^2 = P(\pj{y}|\pj{u_i}) = P(M(\pj{x})|M(\pj{e_i})) = P(x | e_i) = |\<x | e_i \>|^2 \neq 0$.
+Note that we have an arbitrary choice for each $u_i$, since we have to pick a vector from the unit circle (i.e.
+a phase for each basis vector).
+This corresponds to a choice of gauge.
+
+  We also note that the map is colinear, meaning that if $U_X, V_X \subseteq X$ are two subgroups such that $U_X \subset V_X$, then $U_Y, V_Y \subseteq Y$ such that $\pj{U_Y} = M(\pj{U_X})$ and $\pj{V_Y} = M(\pj{V_X})$ are subgroups of $Y$ and $U_Y \subset V_Y$.
+In fact, take a basis $\{e_i\}_{i \in I} \subset X$ such that $\{e_i\}_{i \in I_U \subset I} \subset \{e_i\}_{i \in I_V \subset I}$ are bases for $U_X$ and $V_X$ respectively.
+An element of $X$ belongs to $U_X$ if and only if it is not orthogonal only to elements of the basis of $U_X$ and belongs to $V_X$ only if it not orthogonal only to elements of the basis of $V_X$.
+As the map $M$ preserves orthogonality, these relationships are preserved by the map.
+Therefore $U_Y$ and $V_Y$ are subgroups of $Y$ such that $U_Y \subset V_Y$.
+
+  We now use the gauge freedom to redefine the basis such that for all $i$ we have $M(\pj{e_i}) = \pj{v_i}$ and $M(\pj{e_1 + e_i}) = \pj{v_1 + v_i}$.
+Let $v_1 = u_1$.
+This is the only arbitrary choice we make, and corresponds to the choice of a global phase.
+For each $i>1$, consider $e_1 + e_i$.
+This will belong to the subspace $Sp(e_1, e_i)$.
+This subspace, when mapped through $M$, will give us the subspace spanned by $v_1$ and $u_i$.
+That is, $M(\pj{Sp(e_1, e_i)})=\pj{Sp(v_1, u_i)}$.
+This means we can find a unique $k \in \mathbb{C}$ such that $M(\pj{e_1+e_i})=\pj{v_1+ku_i}$.
+We fix $v_i = k u_i$.
+Note that $P(\pj{e_1} | \pj{e_1+e_i}) = \frac{1}{2}=  P(\pj{e_i} | \pj{e_1+e_i})=P(\pj{v_1} | \pj{v_1+k u_i})=P(\pj{u_i} | \pj{v_1+k u_i})$.
+Therefore $|k| = 1$ and $k u_i = v_i$ is a unit vector.
+
+  Now we want to show that $M(\pj{e_1 + c e_i})=\pj{v_1 + \tau_i(c)
+    v_i}$ where either $\tau_i(c) = c$ or $\tau_i(c) = c^\dagger$. For
+  each $i$, consider $w = e_1 + c e_i \in Sp(e_1, e_i)$. Since
+  $M(\pj{w}) \subset \pj{Sp(v_1, v_i)}$, there must be a $\tau_i(c)$
+  such that $\pj{v_1 + \tau_i(c) v_i} = M(\pj{w})$. Since we must have
+  $P(\pj{e_i} | \pj{w}) = P(\pj{v_i} | M(\pj{w}))$ and $P(\pj{e_1 +
+    e_i} | \pj{w}) = P(\pj{v_1 + v_i} | M(\pj{w}))$, we must have
+  $|c| = |\tau_i(c)|$ and $cos(arg(c)) = cos(arg(\tau_i(c)))$ for any
+  $c$. This means that either $\tau_i(c) = c$ or $\tau_i(c) =
+  c^*$.
+
+  Next we want to show that $\tau_i(c) = \tau_j(c)$ for all pairs $(i,j)$.
+That is, either we have to take the complex conjugate of all components or of none.
+Consider $e_i - e_j$.
+We have $\pj{e_i - e_j} \subset \pj{Sp(e_i, e_j)}$ and, for any $c \in \mathbb{C}$, $\pj{e_i - e_j} \subset \pj{Sp(e_1 + c e_i, e_1 + c e_j)}$.
+By construction, we have $M(\pj{e_i - e_j})\subset \pj{Sp(v_i, v_j)}$ and $M(\pj{e_i - e_j})\subset \pj{Sp(v_1 + \tau_i(c) v_i, v_1 + \tau_j(c) v_j)}$.
+Therefore $M(\pj{e_i - e_j}) = \pj{Sp(v_i, v_j)} \cap \pj{Sp(v_1 + \tau_i(c) v_i, v_1 + \tau_j(c) v_j)} = \pj{\tau_i(c) v_i - \tau_j(c) v_j}$.
+This means that, for all $c$, $\tau_i(c) = \tau_j(c)$.
+
+  Now we show that for all $c_2, ..., c_n \in \mathbb{C}$ we have $M(\pj{e_1 + c_2 e_2 + ...
++ c_n e_n}) = \pj{v_1 + \tau(c_2) v_2 + ...
++ \tau(c_n) v_n}$.
+We prove this by induction.
+If only the first two components are non-zero, we have  $M(\pj{e_1 + c_2 e_2}) = \pj{v_1 + \tau(c_2) v_2}$ by construction.
+Let $2 < p \leq n$.
+If we assume $M(\pj{e_1 + c_2 e_2 + ...
++ c_{p-1} e_{p-1}}) = \pj{v_1 + \tau(c_2) v_2 + ...
++ \tau(c_{p-1}) v_{p-1}}$, then $M(\pj{e_1 + c_2 e_2 + ...
++ c_p e_p}) \subset M(\pj{Sp(e_1 + c_2 e_2 + ...
++ c_{p-1} e_{p-1}, e_p})) = \pj{Sp(v_1 + c_2 v_2 + ...
++ c_{p-1} v_{p-1}, v_p)}$.
+This means that there exists $k_p \in \mathbb{C}$ such that $M(\pj{e_1 + c_2 e_2 + ...
++ c_p e_p}) = \pj{v_1 + c_2 v_2 + ...
++ c_{p-1} v_{p-1} + k_p v_p)}$.
+But we also have $M(\pj{e_1 + c_2 e_2 + ...
++ c_p e_p}) \subset M(\pj{Sp(e_1 + c_p e_p, c_2 e_2 + ...
++ c_{p-1} e_{p-1}})) = \pj{Sp(v_1 + \tau(c_p) v_p, c_2 v_2 + ...
++ c_{p-1} v_{p-1}, v_p)}$.
+The only way this can work is if $k_p = \tau(c_p)$.
+
+  We also need to show the above works when there is no component on the first element of the basis.
+That is, for all $c_2, ..., c_n \in \mathbb{C}$ we have $M(\underbar{$c_2 e_2 + ...
++ c_n e_n$}) = \underbar{$\tau(c_2) v_2 + ...
++ \tau(c_n) v_n$}$.
+First note that $M(\underbar{$c_2 e_2 + ...
++ c_n e_n$}) \subset M(\underbar{$Sp(e_2, ..., e_n)$)} = \underbar{$Sp(v_2, ..., v_n)$}$.
+Also note that $M(\underbar{$c_2 e_2 + ...
++ c_n e_n$}) \subset M(\underbar{$Sp(e_1, e_1 + c_2 e_2 + ...
++ c_n e_n)$)} = \underbar{$Sp(v_1, v_1 + \tau(c_2) v_2 + ...
++ \tau(c_n) v_n)$}$.
+The only way this can work is if $M(\underbar{$c_2 e_2 + ...
++ c_n e_n$}) = \underbar{$\tau(c_2) v_2 + ...
++ \tau(c_n) v_n$}$.
+
+  We can now define $m : X \to Y$ such that $m(e_i) = v_i$ for all $i$ and $m(\sum_{i \in I} c_i e_i) = \sum_{i \in I} \tau(c_i) v_i$.
+This means $\underbar{$m(\sum_{i \in I} c_i e_i)$} = M(\underbar{$\sum_{i \in I} c_i e_i$})$.
+Moreover, if $\tau(c) = c$ we have $\<m(\sum_{i \in I} c_i e_i)|m(\sum_{j \in I} d_j e_j)\> = \<\sum_{i \in I} c_i v_i|\sum_{j \in I} d_j v_j\> = c_i^* d_j \delta_{ij} = \<\sum_{i \in I} c_i e_i|\sum_{j \in I} d_j e_j\>$.
+On the other hand, if $\tau(c) = c^*$ we have $\<m(\sum_{i \in I} c_i e_i)|m(\sum_{j \in I} d_j e_j)\> = \<\sum_{i \in I} c_i^* v_i|\sum_{j \in I} d_j^* v_j\> = \<\sum_{j \in I} d_j v_j|\sum_{i \in I} c_i v_i\> = d_i^* c_j \delta_{ij} = \<\sum_{j \in I} d_j e_j|\sum_{i \in I} c_i e_i\>$.
+\end{proof}
+
+\begin{remark}
+The fact that the proposition identifies either a linear map or an anti-linear (i.e.~conjugate-linear)  has a clear physical interpretation.
+Since, for a Hilbert space, the conjugate vector space is equivalent to the dual space, a conjugate-linear map is one that preserves the inner product but maps ket vectors into bra vectors.
+Looking ahead, the above result does not exclude a composition map similar to the tensor product, but that maps the kets of one or both subsystems into bras in the composite system.
+
+Such a conjugate-linear map is physically indistinguishable from a map that maps kets into kets: the physics is only contained in the probabilities which are the square moduli of inner products.
+The idea is that, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear.
+With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.
+
+Note that this unnecessary subtlety could in principle be avoided by
+reformulating quantum mechanics in terms of quantum states given by
+density matrices $\rho=|\psi\>\<\psi|$ (which contain both  kets and
+bras), as is done, for example in \cite{ozawa,holevo}. In this paper
+we employed the more familiar formulation in which quantum states are
+rays in Hilbert space (identified either by kets or bras).
+\end{remark}
+
+\begin{prop}[Bilinearity, H2]\label{prop_bilinearity}
+  Let $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ be a map such that, for all $(a,b) \in \mathcal{A} \times \mathcal{B}$, we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$.
+Then $m$ must be bilinear.
+That is:
+  \begin{align}
+m(k_1a_1 + k_2a_2, b)=k_1m(a_1, b) + k_2m(a_2, b) \\
+m(a, k_1b_1 + k_2b_2)=k_1m(a, b_1) + k_2m(a, b_2)
+\end{align}
+for all $a, a_1, a_2 \in \mathcal{A}$, $b, b_1, b_2 \in \mathcal{B}$ and $k_1, k_2 \in \mathbb{C}$.
+\end{prop}
+
+\begin{proof}
+  If we fix $b \in \mathcal{B}$, then we have $M_b : \pj{\mathcal{A}} \to \pj{\mathcal{C}}$ where $M_b(\pj{a}) = M(\pj{a}, \pj{b})$.
+By \ref{prop_statInd} and \ref{prop_fundProj} we can find a linear map $m_b : \mathcal{A} \to \mathcal{C}$ such that $\pj{m_b(a)} = M_b(\pj{a}) = M(\pj{a}, \pj{b})$.
+As this must map subspace to subspace, we must have $m(a, b) = k m_b(a)$ for some $k \in \mathbb{C}$.
+Since $m_b$ is linear, we have $m(k_1a_1 + k_2a_2, b)=k_1m(a_1, b) + k_2m(a_2, b)$ for any $a_1, a_2 \in \mathcal{A}$ and $k_1, k_2 \in \mathbb{C}$.
+We can repeat the argument fixing $a \in \mathcal{A}$, and find $m(a, k_1b_1 + k_2b_2)=k_1m(a, b_1) + k_2m(a, b_2)$ for any $b_1, b_2 \in \mathcal{B}$ and $k_1, k_2 \in \mathbb{C}$.
+\end{proof}
+
+\begin{prop}[Subsystems' basis gives composite system
+  basis]\label{prop_basis}
+  Let $\{a_i\}_{i\in I}$ and $\{b_j\}_{j \in J}$ be bases of $\mathcal{A}$ and $\mathcal{B}$ respectively, then a set of unit vectors $\{e_{ij}\}_{(i,j) \in I \times J} \subset \mathcal{C}$  such that $e_{ij} \in M(\underbar{$a_i$}, \underbar{$b_j$})$ forms a basis for $\mathcal{C}$.
+\end{prop}
+
+\begin{proof}
+
+  Since $M$ is a map on the projective spaces, it maps spans to spans.
+  Since the span of the basis of $\mathcal{A}$ and $\mathcal{B}$ is
+  the whole space, then the span of the image of the basis is the
+  whole image of $M$. By \ref{prop_spanSurj}, the image of $M$
+  coincides with the whole $\mathcal{C}$. Therefore, given
+  $\{a_i\}_{i\in I}$ and $\{b_j\}_{j \in J}$ bases of $\mathcal{A}$
+  and $\mathcal{B}$ respectively, any set of unit vectors
+  $\{e_{ij}\}_{(i,j) \in I \times J} \subset \mathcal{C}$ such that
+  $e_{ij} \in M(\pj{a_i}, \pj{b_j})$ spans the whole $\mathcal{C}$.
+
+  Now consider $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))$.
+If $i=k$ and $j=l$ we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))= P(M(\pj{a_i},\pj{b_j})|M(\pj{a_i},\pj{b_j})) = 1$.
+If $i\neq k$, we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l})) = P(\pj{a_i} \wedge \pj{b_j} | \pj{a_k} \wedge \pj{b_l}) \leq P(\pj{a_i} | \pj{a_k} \wedge \pj{b_l})$.
+By \ref{prop_singleBorn} we have $P(\pj{a_i} | \pj{a_k} \wedge \pj{b_l}) = P(\pj{a_i} | \pj{a_k}) = 0$ since $a_i$ and $a_k$ are different elements of an orthogonal basis.
+Therefore we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))=\delta_{ik}\delta_{jl}$ which means $\<e_{ij} | e_{kl} \> = \delta_{ik}\delta_{jl}$.
+
+  The elements $e_{ij}$ form a set of orthonormal vectors that span the whole space and are therefore a basis.
+\end{proof}
+
+\begin{thrm}[Composite system theorem]\label{theo}
+  The state space of a composite system of independent systems is given by the tensor product of the spaces of the component systems.
+\end{thrm}
+
+\begin{proof}
+We are looking for a map $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ such that, for all $(a,b) \in \mathcal{A} \times \mathcal{B}$ we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$.
+We saw in \ref{prop_bilinearity} that if $m$ exists, it must be bilinear.
+
+Now we show that, if $m$ exists, then $\mathcal{C} \cong \mathcal{A}
+\otimes \mathcal{B}$ (where $\cong$ indicates an isomorphism) and $m :
+\mathcal{A} \times \mathcal{B} \to \mathcal{A} \otimes \mathcal{B}$ is
+the standard map from the Cartesian product to the tensor product. As $m
+: \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a bilinear
+operator, by the universal property of the tensor product we can find
+a linear operator $\hat{m} : \mathcal{A} \otimes \mathcal{B} \to
+\mathcal{C}$ such that $m(a, b) = \hat{m}(a \otimes b)$. By
+\ref{prop_basis} the set $\{ m(a_i, b_j)\}_{(i,j) \in I \times J}$
+forms a basis since $ m(a_i, b_j) \in M(\pj{a_i}, \pj{b_j})$ for all
+$(i, j)$, therefore $\hat{m}(\{ a_i \otimes b_j\}_{(i,j) \in I \times
+  J})$ also forms a basis since $\hat{m}(a_i\otimes b_j)=m(a_i, b_j)$.
+By \ref{prop_totality}, each $m(a_i,b_j)$ will correspond to a unit
+vector in $\mathcal{C}$. We have $\<\hat m(a_i\otimes b_j)| \hat
+m(a_k\otimes b_\ell)\>_\mathcal{C} =\<m(a_i, b_j)| m(a_k,
+b_\ell)\>_\mathcal{C} = \delta_{ik}\delta_{j\ell} = \<a_i\otimes b_j|
+a_k \otimes b_\ell\>_{\otimes}$. The function $\hat{m}$, then,
+preserves the inner product across all elements of the basis and is
+therefore an isomorphism for Hilbert spaces. We have $\mathcal{C}
+\cong \mathcal{A} \otimes \mathcal{B}$ and $m(a, b) = \hat{m}(a
+\otimes b) \cong a \otimes b$.
+
+Given that the tensor product map exists and it satisfies all the properties $m$ must satisfy, then $m$ exists and it is the tensor product.
+\end{proof}
+
+To conclude, we give a schematic outline of the logical implications
+that led us to the result. This is an expanded version of Fig.~1 of
+the main paper:
+\begin{enumerate}
+\item P\ref{post_state}: states and observables postulate.
+
+\item P\ref{post_measurements}: Born rule (measurement postulate).
+
+\item Def \ref{def_indep}: Preparation independence: systems are independent if the preparation of one does not affect the other.
+
+\item P\ref{post_measurements} $\Rightarrow$ \ref{prop_singleBorn}: the outcome probabilities depend only on the
+  inner product.
+
+\item Def \ref{def_comp}: Composite system definition: A composite system is a
+  collection of the subsystems (i.e.~all compatible states give a preparation) and only of the subsystems
+  (i.e.~all composite preparations give non-trivial measurements on the
+  subsystems).
+
+\item P\ref{post_measurements} + Def \ref{def_comp} $\Rightarrow$ \ref{prop_spanSurj} (H3): Span
+  surjectivity (all composiste $\cal C$ are superpositions of ${\cal A}$ and ${\cal B}$).
+
+\item Def \ref{def_indep} + Def \ref{def_comp} $\Rightarrow$ \ref{prop_totality} (H1):
+  Totality (all possible state pairs of the subsystems correspond to a state of the composite).
+
+\item P\ref{post_state} + \ref{prop_singleBorn} +  Def \ref{def_comp} $\Rightarrow$ \ref{prop_statInd}: Statistical
+  independence (if one subsystem does not change, the probability on the composite system is given by the probability of the subsystem that changes).
+
+\item \ref{prop_fundProj}: Fundamental theorem of projective geometry (preserving square of inner product leads to unique linear map)
+
+\item \ref{prop_statInd} + \ref{prop_fundProj} $\Rightarrow$ \ref{prop_bilinearity} (H2) composition map on vector spaces is bilinear.
+
+\item P\ref{post_measurements} + Def \ref{def_comp} + \ref{prop_spanSurj} (H3) $\Rightarrow$ \ref{prop_basis}: Basis
+  carries over from subsystems to composite
+
+\item \ref{prop_totality} + \ref{prop_bilinearity} + \ref{prop_basis} $\Rightarrow$ \ref{theo}: the composition map is the tensor product.
 \end{enumerate}
-
-Proposition 1. $C = \otimes $
-
-Proof. Step 1. $C$ is bilinear. Fixing a unitary $b \in \mathcal{B}$, for all $a, a_1, a_2 \in \mathcal{A}$ and $k_1, k_2 \in \mathbb{C}$ we have
-\begin{equation}
-	\begin{aligned}
-	<C(a, b), C(k_1a_1 + k_2a_2, b)>_\mathcal{C} &= <a, k_1a_1 + k_2a_2>_\mathcal{A} \\
-	&= k_1<a, a_1>_\mathcal{A} + k_2<a, a_2>_\mathcal{A} \\
-	&= k_1<C(a, b), C(a_1, b)>_\mathcal{C} + k_2<C(a, b), C(a_2, b)>_\mathcal{C} \\
-	&= <C(a, b), k_1C(a_1, b) + k_2C(a_2, b)>_\mathcal{C} \\
-	\end{aligned}
-\end{equation}
-Since this is valid for all $a, a_1, a_2 \in \mathcal{A}$ and $k_1, k_2 \in \mathbb{C}$ then $C(k_1a_1 + k_2a_2, b)=k_1C(a_1, b) + k_2C(a_2, b)$.
-Similarly, we have $C(a, k_1b_1 + k_2b_2)=k_1C(a, b_1) + k_2C(a, b_2)$.
-$C$ is bilinear.
-
-%Step 1: inner product agrees over all linear operators.
-Let $A^\mathcal{A} : \mathcal{A} \to \mathcal{A}$ be a linear operator.
-We can decompose $A^\mathcal{A} = A_a^\mathcal{A} + \imath A_b^\mathcal{A}$ such that $A_a^\mathcal{A}$ and $A_b^\mathcal{A}$ are Hermitian.
-For all $v_a \in V_\mathcal{A}$, $v_b \in V_\mathcal{B}$ we have:
-%\begin{equation}
-%	\begin{aligned}
-%	<v_a, A^\mathcal{A}v_a>_\mathcal{A} &= <v_a, A_a^\mathcal{A}v_a>_\mathcal{A} + \imath <v_a, A_b^\mathcal{A}v_a>_\mathcal{A} \\
-%	&= <C(v_a, v_b), A_a^\mathcal{C}C(v_a, v_b)>_\mathcal{C} + \imath <C(v_a, v_b), A_b^\mathcal{C}C(v_a, v_b)>_\mathcal{C} \\
-%	&= <C(v_a, v_b), A^\mathcal{C}C(v_a, v_b)>_\mathcal{C}
-%	\end{aligned}
-%\end{equation}
-%Similarly for $\mathcal{B}$.
-
-Step 2: C is the tensor product.
-As $C : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a bilinear operator, for the universal property of the tensor product we can find a linear operator $\hat{C} : \mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that $C(a, b) = \hat{C}(a \otimes b)$.
-The set $E = \{ a\otimes b | <a,a>=1, <b,b>=1 \}$ forms a basis for the tensor product and therefore $\hat{C}(E)$ is a basis for $\mathcal{C}$.
-We have :
-$$<C(a, b), C(a, b)>_\mathcal{C} = <a\otimes b, a \otimes b>_{\otimes} = <a, a>_\mathcal{A} = <b, b>_\mathcal{B} = 1$$
-The function $\hat{C}$, then, is unitary over all elements of the basis and is therefore the identity. We have $C(a, b) = \hat{C}(a \otimes b) = a \otimes b$.
-
-\section{Second try}
-
-Lemma 1. Let $\mathcal{A}$, $\mathcal{B}$ and $\mathcal{C}$ be linear spaces. Let $C : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ be an injective map that satisfy the following:
-\begin{enumerate}
-	\item given a positive definite linear operator $\rho_\mathcal{A} : \mathcal{A} \to \mathcal{A}$, there exists a corresponding positive definite linear operator $\rho_\mathcal{C} : \mathcal{C} \to \mathcal{C}$ such that $\rho_\mathcal{C}(C(a,b)) = C(\rho_\mathcal{A}(a),b)$; moreover $k_1\rho_\mathcal{A}^1 + k_2\rho_\mathcal{A}^2$ corresponds to $k_1\rho_\mathcal{C}^1 + k_2\rho_\mathcal{C}^2$
-	\item given a positive definite linear operator $\rho_\mathcal{B} : \mathcal{B} \to \mathcal{B}$, there exists a corresponding positive definite linear operator $\rho_\mathcal{C} : \mathcal{C} \to \mathcal{C}$ such that $\rho_\mathcal{C}(C(a,b)) = C(a,\rho_\mathcal{B}(b))$; moreover $k_1\rho_\mathcal{B}^1 + k_2\rho_\mathcal{B}^2$ corresponds to $k_1\rho_\mathcal{C}^1 + k_2\rho_\mathcal{C}^2$
-\end{enumerate}
-Then $C$ is a bilinear operator. That is, $C(k_1a_1 + k_2a_2, b)=k_1C(a_1, b) + k_2C(a_2, b)$ and $C(a, k_1b_1 + k_2b_2)=k_1C(a, b_1) + k_2C(a, b_2)$.
-
-Proof.
-\emph{Extend linearity to negative and imaginary operators.} We can univocally extend the correspondence to negative and imaginary operators.
-If $\rho_\mathcal{A}$ is a positive operator, we have $\rho_\mathcal{A} = - (-\rho_\mathcal{A})$.
-If we extend the linear correspondence, we must have $\rho_\mathcal{C} = - (-\rho_\mathcal{A})_\mathcal{C}$ and therefore $(-\rho_\mathcal{A})_\mathcal{C} = - \rho_\mathcal{C}$.
-Similarly, $- \rho_\mathcal{A} = \imath (\imath \rho_\mathcal{A})$ which leads to $(\imath \rho_\mathcal{A})_\mathcal{C} = \imath \rho_\mathcal{C}$.
-One can perform the same extension on $\mathcal{B}$.
-
-\emph{All linear operators have a linear correspondence.} Let $O_\mathcal{A}$ be a linear operator.
-We can always write $O_\mathcal{A} = O_\mathcal{A}^{r+} - O_\mathcal{A}^{r-} + \imath O_\mathcal{A}^{i+} - \imath O_\mathcal{A}^{i-}$ where  $O_\mathcal{A}^{r+}$, $O_\mathcal{A}^{r-}$, $\imath O_\mathcal{A}^{i+}$ and $\imath O_\mathcal{A}^{i-}$ are all positive definite.
-Therefore we can find a corresponding $O_\mathcal{C}$.
-One can perform the same expansion on linear operators on $\mathcal{B}$.
-
-\emph{The map $C$ is bilinear.} Let $a_1, a_2 \in \mathcal{A}$.
-The we can find a linear operator such that $a_2 = O a_1$.
-We have $C(k_1 a_1 + k_2 a_2, b) = C(k_1 I a_1 + k_2 O a_1, b) = (k_1 I + k_2 O)_\mathcal{C}C(a_1, b) = (k_1 I)_\mathcal{C}C(a_1, b) + (k_2 O)_\mathcal{C}C(a_1, b) = C(k_1 I a_1, b) + C(k_2 O a_1, b)  = C(k_1 a, b) + C(k_2 a_2, b)$.
-Similarly, we find $C(a, k_1b_1 + k_2b_2)=k_1C(a, b_1) + k_2C(a, b_2)$.
-The map is bilinear.
-
 \end{document}

### 2020-03-31 Articles/2019-QuantumComposition/Submissions/NatureComm/supplementaryinfo.tex
Preparing Nature Communications submission

--- 2020-03-31 Articles/2019-QuantumComposition/Source/MathCheck.tex
@@
+%ATTENZIONE: PER COMPILARE BISOGNA FARE A MANO
+%dvipdf tensor.dvi ;mv tensor.pdf temp.pdf 
 %Dovremmo forse aggiungere due parole sulle C* algebras: non sono stateintrodotte proprio per evitare il tensor product in infinitedimensional systems?
@@
 
-	\fbox{{\scriptsize Internal report. \today}}\title{The four postulates of quantum mechanics are three}
-	\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^2$ and Christine A. Aidala$^1$
-	}\affiliation{\vbox{1.~Physics Department, University of Michigan, 450 Church Street,
-			Ann Arbor, MI 48109-1040,
-			United States}\\
-		\vbox{2.~Dip.~Fisica and INFN Sez.\ Pavia, University
-			of Pavia, via Bassi 6, I-27100 Pavia, Italy}}
-	\begin{abstract}
-		The tensor product postulate of quantum mechanics states that the
-		Hilbert space of a composite system is the tensor product of the
-		components' Hilbert spaces. All current formalizations of quantum
-		mechanics that do not contain this postulate contain some equivalent
-		postulate or assumption (sometimes hidden). Here we give a natural
-		definition of composite system as a set containing the component
-		systems and show how one can logically derive the tensor product
-		rule from the state postulate and from the measurement postulate. In
-		other words, our paper reduces by one the number of postulates
-		necessary to quantum mechanics.
-	\end{abstract}
-	\pacs{}
-	% Measurement theory (quantum mechanics), 03.65.Ta Mechanics quantum,
-	% 03.65.-w noise quantum, 42.50.Lc quantum information, 03.67.Ac
-	% Quantum information, 03.67.-a Quantum fluctuations, 42.50.Lc quantum
-	% mechanics, 03.65.Ta quantum optics, 42.50.Gy
-	\maketitle
+	\section{Supplementary information for ``The four postulates of quantum
+		mechanics are three''. }\label{app}
+	\par\noindent 
+	Gabriele Carcassi, Lorenzo Maccone, Christine A. Aidala
+	\subsection{ Mathematical formulation.}
+	\setcounter{section}{1}
+	\setcounter{equation}{0}
+	\counterwithout{equation}{section}
+	\renewcommand{\theequation}{S\arabic{equation}}
 
-	The tensor product postulate does not appear in all axiomatizations of
-	quantum mechanics: it has even been called ``postulate 0'' in some
-	literature \cite{zurek}. A widespread belief is that it is a direct
-	consequence of the superposition principle, and it is hence not a
-	necessary axiom. {\em This belief is mistaken}: the superposition
-	principle is encoded into the quantum axioms by requiring that the
-	state space is a {\em linear} vector space. This is, by itself,
-	insufficient to single out the tensor product, as other linear
-	products of linear spaces exist, such as the direct product, the
-	topological product or the direct sum of vector spaces, which are used
-	in classical mechanics to combine state spaces of linear systems.
-	This belief may have arisen from the seminal book of Dirac
-	\cite{diracbook}, who introduces tensor products (Chap.~20) by
-	appealing to linearity. However, he adds the seemingly innocuous
-	request that the product among spaces be distributive (rather,
-	bilinear), which is equivalent to postulating tensor products (or
-	linear functions of them). This is not an innocuous request. For
-	example it does not hold where the composite vector space of two
-	linear spaces is described by the direct product, e.g.~in classical
-	mechanics, for two strings of a guitar: it is not distributive.
-	[General classical systems, not only linear ones, are also composed
-	through the direct product.] Of course, Dirac is not constructing an axiomatic
-	formulation, so his `sleight of hand' can be forgiven. In contrast, von
-	Neumann (\cite{vonneumannbook} Chap.~VI.2, also \cite{jauch})
-	introduces tensor products by noticing that this is a natural choice
-	in the position representation of wave mechanics (where they were
-	introduced in \cite{weyl,epr}), and then {\em explicitly postulates}
-	them in general: ``This rule of transformation is correct in any case
-	for the coordinate and momentum operators [...] and it conforms with
-	the [observable axiom and its linearity principles], we therefore
-	postulate them generally.''  \cite{vonneumannbook}.  More mathematical
-	or conceptually-oriented modern formulations
-	(e.g.~\cite{ozawa,masanes,nielsenchuang}) introduce this postulate explicitly.
-	An interesting alternative is provided in
-	\cite{ballentinebook,ballentinepaper}: after introducing tensor
-	products, Ballentine verifies a posteriori that they give the correct
-	laws of composition of probabilities. Similarly, Peres uses
-	relativistic locality \cite{peres}. While these procedures seemingly
-	bypass the need to postulate the tensor product, they do not guarantee
-	that this is the {\em only} possible way of introducing composite
-	systems in quantum mechanics. In the framework of quantum logic,
-	tensor products arise from some additional conditions \cite{matolcsi}
-	which (in contrast to what is done here) are not connected to the
-	other postulates. A similar approach was followed in \cite{aerts}
-	where tensor products were obtained by specifying some additional
-	physical requirements. In quantum field theory one tends to avoid
-	problems connected with tensor products of infinite dimensional spaces
-	by focusing on algebraic commutation structures, e.g.~\cite{giddins}.
-	In particular, the recent MIP*=RE result \cite{mipre} implies that, in
-	infinite dimensions, the tensor product is strictly less
-	computationally powerful than the commutation structures, emphasizing
-	the difference among these two structures, at least for the
-	infinite-dimensional case. We will consider the non-relativistic
-	setting here.
-
-	In this paper we derive the tensor product postulate (which, hence,
-	loses its status of postulate) from two other postulates of quantum
-	mechanics: the state postulate and the measurement postulate. We start
-	from the natural definition of a composite system as the set of two
-	(or more) quantum systems, in the sense that the composite system is
-	made of system $A$ {\em and} (joined with) system $B$, namely a set
-	whose elements are the two systems and nothing else. We will focus on
-	kinematically-independent systems, namely no superselection rules or
-	other restrictions to the state space are present: it is possible to
-	prepare each subsystem of a composite system in a state that is
-	independent of the other systems (preparation independence).  This is
-	the only case in which the tensor product can be properly employed
-	\cite{susskind,zanardi,zanardilloyd}: the Hilbert space of composite
-	systems that have restrictions is {\em not} the tensor product of the
-	component spaces, but a subspace of it (e.g.~the anti-symmetric
-	subspace for fermions). Typically this is ignored in the literature,
-	since the tensor product formalism is very convenient and is often
-	used also in these cases. Moreover, in accordance with the measurement
-	postulate, we require that the probability distribution of the
-	measurement outcomes of one system is independent of the other systems' (statistical
-	independence). As detailed below, statistical independence is not an
-	additional requirement: it is already contained in the measurement
-	postulate.
-
-	From the above definition of composite system, it follows that there
-	must be a map $M$ that connects the states of the subsystems to the
-	states of the composite system. A quantum state is a ray in Hilbert
-	space, namely a set of vectors. The map $M$ on states 
-	corresponds to a map $m$ on vectors. We then show that preparation
-	independence and statistical independence imply three conditions on
-	the map $m$: (H1)~totality: the map is defined on all states of the
-	subsystems; (H2)~bilinearity: the map is bilinear thanks to the
-	fundamental theorem of projective geometry; (H3)~span surjectivity:
-	the span of the image of map coincides with the full composite Hilbert space.
-	We then prove that, if these three conditions H1, H2 and H3 hold, then
-	the map $m$ is the tensor product, namely the Hilbert space of the
-	composite system is a tensor product of the components: the tensor
-	product ``postulate'', which hence loses its status of a postulate. An
-	overview of all these logical implications is given in
-	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
-	argument. Refer to the supplementary material for a more rigorous proof of the \emph{same} argument.
-
-	\begin{figure}[h]
-		\includegraphics[width=\linewidth, trim={0.2in 5.2in 2.5in 0.6in}, clip=true]{fig.eps}
-		\caption{Schematic depiction of the logical implications used here. FTPG stands for ``Fundamental Theorem of Projective Geometry''
-			\label{f:fig}}\end{figure}
-	\togli{
-		\begin{widetext}\begin{verbatim}
-			* P1: states and observable postulate.
-			* P2: Born rule (measurement postulate). 
-			* P1 -> systems can be prepared unconditionally in any state.
-			* P2 -> the outcome probabilities depend only on the scalar
-			product.
-			* 2. Def: Composite system definition: A composite system is a collection of
-			(all compatible state give a 
-			preparation) and only of (all composite preparations give non-trivial
-			measurement on subsystems) the subsystems 
-			* P1+ Def -> 1. Preparation independence (all state pairs are compatible)
-			* P2+ Def -> 3. Span surjectivity (I.4=H3) (all C is spanned by AxB) 
-			* 3 -> 6. Basis carries over I.7
-			* 1 + 2 -> 4. Injectivity (M total function) I.5=H1
-			* 2 + PI -> 5. Statistical independence I.6 (probability on compatible
-			pairs where one subsystem does not change) 
-			* 5 -> 7. I.8=H2: scalar product carries over (fundamental theorem of
-			projective geometry).
-			* 7 + 6 + 4 (i.e. H1+H2+H3) -> 8. m is tensor product   
-			\end{verbatim}\end{widetext}
-	}
-
-	We start by noting that the tensor product is uniquely
-	characterized, up to isomorphism, by a universal property regarding
-	bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
-	tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
-	map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the property
-	than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ factors
-	through $T$ uniquely.  This means that there exists a {\em unique}
-	$\hat m$, dependent on $m$, such that $\hat m \circ T=m$.  In other
-	words, the following diagram commutes:
-	\begin{center}
-		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{m}"] \\
-			& \mathcal{C}
-		\end{tikzcd}
-	\end{center}
-
-	We use the axiomatization of quantum mechanics based on the following
-	postulates (e.g.~\cite{ozawa,nielsenchuang}): (a)~The state of a
-	system is described by a ray $\pj{\psi}$ corresponding to a set of
-	non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
-	system's observable properties are described by self-adjoint operators
-	acting on that space; (b)~The probability that a measurement of a
-	property $X$, described by the operator with spectral decomposition
-	$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
-	$x$ given that the system is in state $\pj{\psi}$ is
-	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2$ (Born rule). (c)~The state
-	space of a composite system is given by the tensor product of the
-	spaces of the component systems; (d)~The time evolution of an isolated
-	system is described by a unitary operator acting on a vector
-	representing the system state, $|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or,
-	equivalently, by the Schr\"odinger equation. The rest of quantum
-	theory can be derived from these axioms. While some axiomatizations
-	introduce further postulates, we will be using only (a) and (b) to
-	derive (c), so the above are sufficient for our aims.
-
-	\togli{This axiomatization implicitly contains a definition of ``quantum
-		system'' which is crucial for what follows, so we need to clarify the
-		assumptions that it contains. We will use the following definition for
-		a quantum system\togli{$\stackon[1pt]={\mbox{\tiny
-					def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree of
-			freedom with $d$ (possibly discrete, or continuous, infinite)
-			mutually exclusive (commuting) values for each of its properties.
-			Its mathematical description is through a Hilbert space of dimension
-			$d$ which contains all the states that describe the values of its
-			possible properties. In accordance with the postulate (a), these
-			values correspond to a basis of the space, given by the eigenvectors
-			of the observable corresponding to that property''}. \commentg{We may have to revise to be more clear. Where is this used?} }
-	As stated
-	above, we will limit ourselves to kinematically-independent systems,
-	where all state vectors $|\psi\>$ in the system's Hilbert space $\cal H$
-	describe a valid state, {\em unconditioned on anything else}. In
-	particular, we note that restrictions due to superselection rules
-	arise either from practical (not fundamental) limitations on the
-	actions of the experimenter \cite{susskind,zanardi,zanardilloyd} or
-	from the use of ill-defined quantum systems. For example, in
-	situations where indistinguishability plays a role (e.g.~in QFT), one
-	cannot consider an electron as a quantum system: in this case the
-	quantum system is the field. The electron is an excitation of the
-	field, {\em not} a (sub)system.
-
-	This implies that the ``system'' is kinematically independent from
-	anything else.  We call this condition ``preparation independence''.
-	[We emphasize that the kinematic independence is inequivalent to
-	dynamical independence (or isolation).  Indeed if two systems
-	interact, their interaction may lead to dynamical restrictions in the
-	state spaces. We will not consider dynamical evolution in this paper,
-	which is contained in postulate (d).]
-
-	The definition of a composite system as containing {\em only} the
-	collection of the subsystems means that any preparation of both
-	subsystems independently must correspond to the preparation of the
-	composite system. Since states are defined by postulate (a) as rays in
-	the respective Hilbert spaces, there must exist a map $M : \pj{\cal A}
-	\times \pj{\cal B } \to \pj{\cal C}$ that takes a pair of states for
-	the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$ represent the
-	projective spaces and the Cartesian product is the set of all possible
-	pairs) and returns a state in the projective space $\pj{\cal C}$ for
-	the composite. The map $M$ that acts on rays corresponds to a map $m:{\cal A}\times{\cal B}\to{\cal C}$ that
-	acts on vectors in the Hilbert spaces $\cal A$, $\cal B$ and $\cal C$.
-	Namely, $\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign
-	indicates the elements in the projective space. We will prove that the
-	map $m$ is the tensor product.
-
-	The map $M$ must be injective: as said above, different states of the
-	subsystems must correspond, by definition of composite system, to
-	different states of the composite. Moreover, preparation independence
-	implies that $M$, and hence $m$, must be total maps: each subsystem of
-	the composite system can be independently prepared and gives rise to a
-	state of the composite (condition H1).  H1 is not sufficient
-	to identify the tensor product: by itself it does not even guarantee
-	that the map $m$ is linear.
-
-	Postulate (b) contains the connection between quantum mechanics
-	and probability theory. It must then implicitly contain the
-	axiomatization of probability, e.g.~see
-	\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
-	probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
-	the joint probability of events $a$ and $b$ given $z$ is
-	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Then the events $a$ and $b$ are
-	independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$ (this
-	is the {\em definition} of independent events). Since the Born
-	probability formula in postulate (b) contains only quantities of the
-	system (the state and the observable's eigenstates {\em of the
-		system}), it implies the ``statistical independence'' between
-	different systems if they are prepared independently. Namely, the
-	outcome probability of one is independent of any properties of any
-	other in the absence of any dynamical coupling and supposing
-	independent preparation. [Of course, the dynamics may couple the state
-	of the system to other systems so that the outcomes may depend on what
-	happens to other systems.  Analogously, one may prepare the system in
-	a way that depends on other systems. But we consider independent
-	preparations and do not consider dynamics, as it is sufficient for our
-	aims.] We can then formalize ``statistical independence'' of
-	independently prepared systems by saying that their Born rule
-	satisfies:
-	\begin{eqnarray}
-	p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
-		B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
-		A})\:p(b|\psi_{\cal B})
-	\labell{bindip}\;,
-	\end{eqnarray}
-	where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
-	preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
-	$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
-	B$ respectively. The first equality in \eqref{bindip} embodies the
-	definition of statistical independence, the second follows from the
-	form of the Born rule for independently prepared systems (the
-	probability depends only on the state and observables of each system
-	on its own).
-
-	A particular case is when one system is
-	prepared in a fixed eigenstate $|b\>$ of the observable measured on it,
-	$p(a\wedge b|\psi\wedge b)=p(a|\psi)\:p(b|b)=p(a|\psi)$, where the last equality follows from the Born rule since $\<b|b\>=1$. 
-	If we define $M_b(\pj{a}) = M(\pj{a},\pj{b})$, substituting the values of the
-	probabilities from the Born rule, we have:
-	\begin{align} &\Big|\Big\<M\left(\pj{a},\pj{b}\right)\Big|M\left(\pj{\psi},\pj{b}\right)\Big\>_{\cal C}\Big|^2
-	=\Big|\Big\<M_b\left(\pj{a}\right)\Big|M_b\left(\pj{\psi}\right)\Big\>_{\cal C}\Big|^2
-	\nonumber \\&
-	=\left|\<a|\psi\>_{\cal A}\right|^2
-	\labell{questa},
-	\end{align}
-	where the first and second terms contain the scalar product in the composite
-	space $\cal C$. [This is not a new assumption: it follows from the
-	measurement postulate (b) for the composite system.] This means that,
-	when one subsystem is prepared in an eigenstate of what is measured
-	there, the state space of the other is mapped preserving the square of
-	the inner product.
-
-	Since the square of the inner product is preserved, orthogonality and
-	the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
-	colinear transformation by definition. In this case, the fundamental
-	theorem of projective geometry \cite{fun} applies, which tells us that
-	a unique semi-linear map $m_b$ that acts on the vectors exists in accordance with $M_b$.
-	Moreover, conservation of probability further constrains it to be
-	either linear or antilinear. This tells us that the corresponding $m$
-	is either linear or antilinear in the first argument. Namely, if
-	\eqref{questa} holds, then
-	\begin{align}
-	\<a|\psi\>=\<m(a,b)|m(\psi,b)\>\labell{h2}\;
-	\\\mbox{ or }
-	\<a|\psi\>=\<m(\psi,b)|m(a,b)\> \labell{h2b}.
-	\end{align}
-	We can ignore the antilinear case \eqref{h2b}, which simply
-	corresponds to the map $m$ mapping kets of one or both subsystems into
-	bras before outputting the composite system (all these maps are
-	trivially physically equivalent, since the dual-space description of a
-	quantum system through bras is equivalent to the description using
-	kets). We can now repeat the same analysis for the second argument of
-	$m$ to conclude that it is a bilinear map, condition (H2).
-
-	The last condition (H3) follows directly from the definition of a
-	composite system. Since it is composed {\em only} of the component
-	systems, for any state $c$ of the composite system, we must find at least one pair $(a, b)$ such that $p(a\wedge b | c)\neq 0$. It follows that the map $m$ is span-surjective: namely the
-	span of the map applied to all states in the component systems spans
-	the composite system state space. In other words, the composite does
-	not contain states that are totally independent of (i.e.~orthogonal
-	to) the states of the components.
-
-	We have obtained the conditions H1, H2 and H3 from the state postulate
-	(a), the measurement postulate (b) and the definition of composite
-	systems. We now prove that these three conditions imply that the (up
-	to now unspecified) composition rule $m$ is the tensor product. More
-	precisely, given a total, span-surjective, bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ that preserves the square of the inner
-	product, we find that $\cal C $ is equivalent to
-	$\cal A \otimes \cal B $ and that $m=\otimes$.
-
-	Proof. Step 1: the bases of the component systems are mapped to a
-	basis of the composite system. Because of totality property (H1) and
-	because the square of the inner product is preserved, we can conclude
-	that, given two orthonormal bases $\{|a_i\>\}\in{\cal A}$ and
-	$\{|b_j\>\}\in{\cal B}$,
-	$|\<m(a_i,b_j)|m(a_k,b_\ell)\>|^2=\delta_{ik}\delta_{j\ell}$, namely
-	$\{|m(a_i,b_j)\>\}$ is an orthonormal set in $\cal C$.  Moreover, the
-	surjectivity property (H3) guarantees that in $\cal C$ no vectors are
-	orthogonal to this set. This implies that it is a basis for $\cal C$.
-
-	Step 2: Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is
-	a bilinear operator (property H2), thanks to the universal property of
-	the tensor product we can find a unique linear operator $\hat{m} :
-	\mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that $m(a, b) =
-	\hat m(a \otimes b)$. The set $\{ \hat m(a_i\otimes b_j)$ with
-	$|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and ${\cal B}\}$
-	forms a basis for $\cal C$, since $\hat m(a_i\otimes b_j)=m(a_i,b_j)$
-	and we have shown above that the latter is a basis.  Thus,
-	\begin{align} 
-	&\<\hat m(a_i\otimes b_j)|\hat m(a_k\otimes b_\ell)
-	\>_{\mathcal{C}}=\<m(a_i,
-	b_j)|m(a_k,b_\ell)\>_\mathcal{C} \nonumber\\& =
-	\delta_{ik}\delta_{j\ell}
-	= \<a_i\otimes
-	b_j| a_k \otimes b_\ell\>_{\otimes},
-	\labell{ecco}\; 
-	\end{align}
-	where we used the orthonormality of the bases and the fact that
-	$|a_i\otimes b_j\>$ is a basis of the tensor product space ${\cal
-		A}\otimes{\cal B}$. The function $\hat{m}$, then, is an isomorphism
-	over all elements of the basis $|a_i\otimes b_j\>$. It is then an
-	isomorphism between $\cal C$ and ${\cal A}\otimes{\cal B}$. Namely,
-	$|m(a_i,b_j)\>_{\cal C} = |\hat{m}(a_i\otimes b_j)\>_{\cal C} \cong
-	|a_i\otimes b_j\>_\otimes$, where the first two vectors (in $\cal C$)
-	are isomorphic to the last vector (in the tensor product space ${\cal
-		A}\otimes{\cal B}$). Namely, $\hat{m}$ maps bijectively to the
-	tensor product, which implies the isomorphism ${\cal C}\cong{\cal
-		A}\otimes{\cal B}$, i.e.~their equivalence through the map
-	$\hat{m}$. Moreover, since ${\cal C}\cong{\cal A}\otimes{\cal B}$,
-	then the map $m : \cal A \times \cal B \to \cal C$ is equivalent to
-	the map $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$,
-	i.e.~the maps $m$ and $\otimes$ are also equivalent through $\hat
-	m$.$\square$
-
-	\vskip 1\baselineskip A few comments on the proof: it is based on the
-	universal property of the tensor product, which uniquely characterizes
-	it. In Step 1 we showed that the bilinear map $m$ maps subsystems'
-	bases into the composite system basis. We also know that there exists a
-	tensor product map $ {T}=\otimes$ that can compose the vectors in $\cal A$
-	and $\cal B$. In Step 2 we use the universal property: since $m$ is a
-	bilinear map, we are assured that there exists a unique $\hat m$ such
-	that $\hat m \circ T=m$. Since we show that $\hat m$ is an
-	isomorphism, then $\hat{m}$ bijectively maps vectors in $\cal C$
-	onto vectors in the tensor product space. Namely $m={T}=\otimes$.
-
-	We conclude with some general comments. The tensor product structure
-	of quantum systems is not absolute, but depends on the observables
-	that are accessible \cite{zanardi,zanardilloyd}. This is due to the
-	fact that an agent that has access to a set of observables will define
-	quantum systems differently from an agent that has access to a
-	different set of observables. Where one agent sees a single system, an
-	agent that has access to less refined observables (and is then limited
-	by some superselection rules) can consider the same system as composed
-	of multiple subsystems. A typical example \cite{tellerbook} comes from
-	quantum field theory. It is customary in basically all quantum optics
-	literature to treat different modes of the radiation field (e.g.~the
-	output of two lasers) as independent systems composed through the
-	tensor product.  Clearly the electromagnetic field is a single system
-	and an agent who is able to access an optical mode that is a linear
-	combination of the two will give a quantum description for it that
-	cannot easily accommodate tensor products. Similarly, an agent can
-	consider two electrons as two systems, joined with the tensor product,
-	whenever they are distinguishable for all practical purposes (e.g.~the
-	electrons are in widely separated physical locations). Yet, in
-	principle, electrons are just excitations of a field, and the `true'
-	quantum system is the field, not the single electrons
-	\cite{teller,tellerbook}.  So, in quantum field theory, the quantum
-	systems that should be joined through tensor products are the
-	different fields and {\em not} the particles, which are just
-	excitations (states) of the fields. In the words of Teller
-	(\cite{tellerbook}, pg.22), tensor products can be safely used only if
-	there is a ``primitive thisness'', which is captured in the definition
-	of system.\togli{\comment{forse e' da spiegare meglio}}
-
-	\togli{\comment{Commento finale da aggiungere?  E' interessante perche' tutto
-			cio' ci dice che non otteniamo necessariamente il prodotto tensore,
-			ma il prodotto tensore modulo una fase locale che e' fisicamente
-			irrilevante. Viene da chiedersi se c'e' una qualche notazione
-			(evidentemente piu' generale di quella di Dirac) che ci permetta di
-			eliminare questa ambiguita' fisicamente irrilevante... La notazione
-			di Dirac gia' elimina la necessita' di avere una rappresentazione
-			per descrivere gli stati. Pero' evidentemente non elimina del tutto
-			la ambiguita' di rappresentazione perche' un cambio di coordinate
-			sia sugli stati che sugli operatori mi lascia invariata la fisica,
-			ma non lascia invariata la notazione di Dirac...  Probabilmente la
-			notazione che elimina l'ambiguita' e' quella che utilizza le matrici
-			densita' invece dei vettori di stato (vedi Ozawa \cite{ozawa} e
-			Holevo): le matrici densita' normalizzate sono phase-independent:
-			$\rho=|\psi\>\<\psi|$. Attenzione: l'interpretazione di una matrice
-			mista $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$ come ``il sistema e' nello
-			stato $|\psi_i\>$ con probabilita' $p_i$'' e' una CONSEGUENZA della
-			regola di Born, quindi nella formalizzazione del postulato degli
-			stati in termini di matrici densita', questa interpretazione non
-			puo' apparire perche' e' una conseguenza di un altro postulato.}}
-
-	It has been pointed out before that the quantum postulates are
-	redundant: in \cite{masanes} it was shown that the measurement
-	postulate (b) can be derived from the others (a), (c), (d). Here
-	instead we have shown how the tensor product postulate can be
-	logically derived from the state postulate (a), the measurement
-	postulate (b) and a reasonable definition of independent systems, and
-	we have described the logical relations among them.  Of course, we do
-	not claim that this is the {\em only} way to obtain the tensor product
-	postulate from the others.
-
-	L.M. acknowledges useful discussions with M.~Ozawa, P.~Zanardi,
-	S.~Lloyd, D.~Zeh, G.~Auletta, A.~Aldeni. We acknowledge funding from
-	the Attract project through the Eu Horizon 2020 research and
-	innovation programme under grant agreement No 777222.
-	\begin{references}
-		\bibitem{zurek} W.H. Zurek, Quantum Darwinism, Nature Phys. {\bf 5},
-		181 (2009).
-		\bibitem{diracbook}P.A.M. Dirac, The principles of quantum mechanics,
-		(Clarendon Press, Oxford, 1966).
-		\bibitem{vonneumannbook}J. von Neumann, Mathematical Foundations of
-		Quantum Mechanics (Princeton Univ.  Press, 1955).
-		\bibitem{jauch}J.M. Jauch, Foundations of quantum mechanics
-		(Addison-Welsey, 1968), pg.~176.
-		\bibitem{weyl} H. Weyl, Gruppentheorie und Quantenmechanik (Hirzel,
-		Leipzig, 1928); translated by H. P. Robertson, The Theory of Groups
-		and Quantum Mechanics (Methuen, London, 1931); reprinted by Dover,
-		p. 91.
-		\bibitem{epr}A. Einstein, B. Podolsky, N. Rosen, Can
-		quantum-mechanical description of physical reality be considered
-		complete?, Phys. Rev. {\bf 47}, 777 (1935).
-		\bibitem{ozawa}M. Ozawa, {Uncertainty relations for noise and
-			disturbance in generalized quantum measurements}, Ann. Phys.  {\bf
-			311}, 350 (2004).
-		\bibitem{masanes}L. Masanes, T.D. Galley, M.P. M\" uller, The
-		measurement postulates of quantum mechanics are operationally
-		redundant, Nat. Commun. {\bf 10}, 1361 (2019).
-		\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum Computation
-		and Quantum Information (Cambridge University Press, Cambridge,
-		2000).
-		\bibitem{ballentinebook}L.E. Ballentine, Quantum Mechanics, a modern
-		development (World Scientific, 2014).
-		\bibitem{ballentinepaper}L.E. Ballentine, Probability theory in
-		quantum mechanics, Am. J. Phys. {\bf 54}, 883 (1986).
-		\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
-		press, 1961).
-		\bibitem{peres}A.~Peres, Classical interventions in quantum systems.
-		II. Relativistic invariance, Phys. Rev. A {\bf 61}, 022117 (2000).
-		\bibitem{matolcsi} T. Matolcsi, Tensor product of Hilbert lattices and
-		free orthodistributive product of orthomodular lattices, Acta Sci.
-		Math. (Szeged), {\bf 37}, 263 (1975).
-		\bibitem{aerts} D. Aerts, I. Daubechies, Physical justification for
-		using the tensor product to describe two quantum systems as one
-		joint system, Helv. Phys. Acta {\bf 51}, 661 (1979).
-		\bibitem{giddins}S.B.~Giddings, Hilbert space structure in quantum
-		gravity: an algebraic perspective. J. High Energ. Phys. 2015, 1
-		(2015).% https://doi.org/10.1007/JHEP12(2015)099
-		\bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
-		rule, Phys. Rev. {\bf 155}, 1428 (1967).
-		\bibitem{zanardi}P. Zanardi, Virtual Quantum Subsystems, Phys. Rev.
-		Lett {\bf 87}, 077901 (2001).
-		\bibitem{zanardilloyd} P. Zanardi, D.A. Lidar, S. Lloyd, Quantum
-		Tensor Product Structures are Observable Induced, Phys. Rev. Lett.
-		{\bf 92},060402 (2004).
-		\bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
-		Field Theory (Princeton Univ. Press, 1997).  
-		\bibitem{teller}M. Redhead, P. Teller, Particles, Particle Labels, and
-		Quanta: The Toll of Unacknowledged Metaphysics, Found. Phys. {\bf
-			21}, 43 (1991).
-		\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
-		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
-		\bibitem{holevo}A. Holevo, Probabilistic and statistical aspects of
-		quantum theory, (North Holland, 1982).
-	\end{references}
-
-	\vskip 5\baselineskip
-
-	\section{Appendix: mathematical formulation}\label{app}
-	\setcounter{section}{1}
+	Here we give the mathematical details of the proof sketched in the
+	main paper.
 
@@
 
-		We also note that the map is colinear, meaning that if $U_X, V_X \subseteq X$ are two subgroups such that $U_X \subset V_X$, then $U_Y, V_Y \subseteq Y$ such that $\pj{U_Y} = M(\pj{U_X})$ and $\pj{V_Y} = M(\pj{V_X})$ are subgroups of $Y$ and $U_Y \subset V_Y$.
-In fact, take a basis $\{e_i\}_{i \in I} \subset X$ such that $\{e_i\}_{i \in I_U \subset I} \subset \{e_i\}_{i \in I_V \subset I}$ are bases for $U_X$ and $V_X$ respectively.
+		We also note that the map is colinear, meaning that if $U_X, V_X \subseteq X$ are two subspaces such that $U_X \subset V_X$, then $U_Y, V_Y \subseteq Y$ such that $\pj{U_Y} = M(\pj{U_X})$ and $\pj{V_Y} = M(\pj{V_X})$ are subspaces of $Y$ and $U_Y \subset V_Y$.
+In fact, take an orthonormal basis $\{e_i\}_{i \in I} \subset X$ such that $\{e_i\}_{i \in I_U \subset I} \subset \{e_i\}_{i \in I_V \subset I}$ are bases for $U_X$ and $V_X$ respectively.
 An element of $X$ belongs to $U_X$ if and only if it is not orthogonal only to elements of the basis of $U_X$ and belongs to $V_X$ only if it not orthogonal only to elements of the basis of $V_X$.
 As the map $M$ preserves orthogonality, these relationships are preserved by the map.
-Therefore $U_Y$ and $V_Y$ are subgroups of $Y$ such that $U_Y \subset V_Y$.
+Therefore $U_Y$ and $V_Y$ are subspaces of $Y$ such that $U_Y \subset V_Y$.
+
+		Additionally we note that, for any colinear map, given two subspaces $U_1, U_2 \subseteq X$ we have $M(\pj{Sp(U_1, U_2)}) = \pj{Sp(M(\pj{U_1}), M(\pj{U_2}))}$.
+In fact, $\pj{Sp(U_1,U_2)}$ is the smallest subspace containing all vectors in $\pj{U_1}$ and $\pj{U_2}$.
+In the same way, $\pj{Sp(M(U_1),M(U_2))}$ is the smallest subspace containing all vectors in $M(\pj{U_1})$ and $M(\pj{U_2})$.
+Since the subspace inclusion is preserved by $M$, we must have $M(\pj{Sp(U_1, U_2)}) = \pj{Sp(M(\pj{U_1}), M(\pj{U_2}))}$.
 
@@
 		\mathcal{A} \times \mathcal{B} \to \mathcal{A} \otimes \mathcal{B}$ is
-		the standard map from the scalar product to the tensor product. As $m
+		the standard map from the Cartesian product to the tensor product. As $m
 		: \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a bilinear
@@
 	\end{enumerate}
+
+	\begin{references}
+		\bibitem{zurek} W.H. Zurek, Quantum Darwinism, Nature Phys. {\bf 5},
+		181 (2009).
+		\bibitem{diracbook}P.A.M. Dirac, The principles of quantum mechanics,
+		(Clarendon Press, Oxford, 1966).
+		\bibitem{vonneumannbook}J. von Neumann, Mathematical Foundations of
+		Quantum Mechanics (Princeton Univ.  Press, 1955).
+		\bibitem{jauch}J.M. Jauch, Foundations of quantum mechanics
+		(Addison-Welsey, 1968), pg.~176.
+		\bibitem{weyl} H. Weyl, Gruppentheorie und Quantenmechanik (Hirzel,
+		Leipzig, 1928); translated by H. P. Robertson, The Theory of Groups
+		and Quantum Mechanics (Methuen, London, 1931); reprinted by Dover,
+		p. 91.
+		\bibitem{epr}A. Einstein, B. Podolsky, N. Rosen, Can
+		quantum-mechanical description of physical reality be considered
+		complete?, Phys. Rev. {\bf 47}, 777 (1935).
+		\bibitem{ozawa}M. Ozawa, {Uncertainty relations for noise and
+			disturbance in generalized quantum measurements}, Ann. Phys.  {\bf
+			311}, 350 (2004).
+		\bibitem{masanes}L. Masanes, T.D. Galley, M.P. M\" uller, The
+		measurement postulates of quantum mechanics are operationally
+		redundant, Nat. Commun. {\bf 10}, 1361 (2019).
+		\bibitem{wootters}W.K. Wootters, Optimal Information Transfer and
+		Real-Vector-Space Quantum Theory. In: Chiribella G., Spekkens R.
+		(eds) Quantum Theory: Informational Foundations and Foils,
+		Fundamental Theories of Physics, vol 181. Springer, Dordrecht
+		(2016).
+		\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum Computation
+		and Quantum Information (Cambridge University Press, Cambridge,
+		2000).
+		\bibitem{ballentinebook}L.E. Ballentine, Quantum Mechanics, a modern
+		development (World Scientific, 2014).
+		\bibitem{ballentinepaper}L.E. Ballentine, Probability theory in
+		quantum mechanics, Am. J. Phys. {\bf 54}, 883 (1986).
+		\bibitem{peres}A.~Peres, Classical interventions in quantum systems.
+		II. Relativistic invariance, Phys. Rev. A {\bf 61}, 022117 (2000).
+		\bibitem{matolcsi} T. Matolcsi, Tensor product of Hilbert lattices and
+		free orthodistributive product of orthomodular lattices, Acta Sci.
+		Math. (Szeged), {\bf 37}, 263 (1975).
+		\bibitem{aerts} D. Aerts, I. Daubechies, Physical justification for
+		using the tensor product to describe two quantum systems as one
+		joint system, Helv. Phys. Acta {\bf 51}, 661 (1979).
+		\bibitem{giddins}S.B.~Giddings, Hilbert space structure in quantum
+		gravity: an algebraic perspective. J. High Energ. Phys. 2015, 1
+		(2015).% https://doi.org/10.1007/JHEP12(2015)099
+		\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
+		\bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
+		rule, Phys. Rev. {\bf 155}, 1428 (1967).
+		\bibitem{zanardi}P. Zanardi, Virtual Quantum Subsystems, Phys. Rev.
+		Lett {\bf 87}, 077901 (2001).
+		\bibitem{zanardilloyd} P. Zanardi, D.A. Lidar, S. Lloyd, Quantum
+		Tensor Product Structures are Observable Induced, Phys. Rev. Lett.
+		{\bf 92},060402 (2004).
+		\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
+		press, 1961).
+		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
+		\bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
+		Field Theory (Princeton Univ. Press, 1997).  
+		\bibitem{teller}M. Redhead, P. Teller, Particles, Particle Labels, and
+		Quanta: The Toll of Unacknowledged Metaphysics, Found. Phys. {\bf
+			21}, 43 (1991).
+		\bibitem{holevo}A. Holevo, Probabilistic and statistical aspects of
+		quantum theory, (North Holland, 1982).
+	\end{references}
+
+	\vskip 1\baselineskip
+
 \end{document}

### 2020-03-31 Articles/2019-QuantumComposition/Submissions/NatureComm/tensor.natcom.tex
Preparing Nature Communications submission

--- 2020-03-31 Articles/2019-QuantumComposition/Source/MathCheck.tex
@@
 
-	\fbox{{\scriptsize Internal report. \today}}\title{The four postulates of quantum mechanics are three}
-	\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^2$ and Christine A. Aidala$^1$
+%\fbox{{\scriptsize Preprint. \today}}
+\title{The four postulates of quantum mechanics are three}
+\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^{2,*}$ and Christine A. Aidala$^1$
 }\affiliation{\vbox{1.~Physics Department, University of Michigan, 450 Church Street,
@@
 [General classical systems, not only linear ones, are also composed
-	through the direct product.] Of course, Dirac is not constructing an axiomatic
-	formulation, so his `sleight of hand' can be forgiven. In contrast, von
-	Neumann (\cite{vonneumannbook} Chap.~VI.2, also \cite{jauch})
-	introduces tensor products by noticing that this is a natural choice
-	in the position representation of wave mechanics (where they were
-	introduced in \cite{weyl,epr}), and then {\em explicitly postulates}
-	them in general: ``This rule of transformation is correct in any case
-	for the coordinate and momentum operators [...] and it conforms with
-	the [observable axiom and its linearity principles], we therefore
-	postulate them generally.''  \cite{vonneumannbook}.  More mathematical
-	or conceptually-oriented modern formulations
-	(e.g.~\cite{ozawa,masanes,nielsenchuang}) introduce this postulate explicitly.
-	An interesting alternative is provided in
-	\cite{ballentinebook,ballentinepaper}: after introducing tensor
-	products, Ballentine verifies a posteriori that they give the correct
-	laws of composition of probabilities. Similarly, Peres uses
+through the direct product.] Of course, Dirac is not constructing an
+axiomatic formulation, so his `sleight of hand' can be forgiven. In
+contrast, von Neumann (\cite{vonneumannbook} Chap.~VI.2, also
+\cite{jauch}) introduces tensor products by noticing that this is a
+natural choice in the position representation of wave mechanics (where
+they were introduced in \cite{weyl,epr}), and then {\em explicitly
+  postulates} them in general: ``This rule of transformation is
+correct in any case for the coordinate and momentum operators [...]
+and it conforms with the [observable axiom and its linearity
+principles], we therefore postulate them generally.''
+\cite{vonneumannbook}.  More mathematical or conceptually-oriented
+modern formulations (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang})
+introduce this postulate explicitly.  An interesting alternative is
+provided in \cite{ballentinebook,ballentinepaper}: after introducing
+tensor products, Ballentine verifies a posteriori that they give the
+correct laws of composition of probabilities. Similarly, Peres uses
 relativistic locality \cite{peres}. While these procedures seemingly
@@
 states of the composite system. A quantum state is a ray in Hilbert
-	space, namely a set of vectors. The map $M$ on states 
-	corresponds to a map $m$ on vectors. We then show that preparation
-	independence and statistical independence imply three conditions on
-	the map $m$: (H1)~totality: the map is defined on all states of the
-	subsystems; (H2)~bilinearity: the map is bilinear thanks to the
-	fundamental theorem of projective geometry; (H3)~span surjectivity:
-	the span of the image of map coincides with the full composite Hilbert space.
-	We then prove that, if these three conditions H1, H2 and H3 hold, then
+space, namely a set of vectors. The map $M$ on states corresponds to a
+map $m$ on vectors. We then show that preparation independence and
+statistical independence imply three conditions on the map $m$:
+(H1)~totality: the map is defined on all states of the subsystems;
+(H2)~bilinearity: the map is bilinear thanks to the fundamental
+theorem of projective geometry; (H3)~span surjectivity: the span of
+the image of map coincides with the full composite Hilbert space.  We
+then prove that, if these three conditions H1, H2 and H3 hold, then
 the map $m$ is the tensor product, namely the Hilbert space of the
@@
 Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
-	argument. Refer to the supplementary material for a more rigorous proof of the \emph{same} argument.
+argument. Refer to the Supplementary Information for a more rigorous
+proof of the \emph{same} argument.
 
-	\begin{figure}[h]
-		\includegraphics[width=\linewidth, trim={0.2in 5.2in 2.5in 0.6in}, clip=true]{fig.eps}
-		\caption{Schematic depiction of the logical implications used here. FTPG stands for ``Fundamental Theorem of Projective Geometry''
-			\label{f:fig}}\end{figure}
-	\togli{
-		\begin{widetext}\begin{verbatim}
-			* P1: states and observable postulate.
-			* P2: Born rule (measurement postulate). 
-			* P1 -> systems can be prepared unconditionally in any state.
-			* P2 -> the outcome probabilities depend only on the scalar
-			product.
-			* 2. Def: Composite system definition: A composite system is a collection of
-			(all compatible state give a 
-			preparation) and only of (all composite preparations give non-trivial
-			measurement on subsystems) the subsystems 
-			* P1+ Def -> 1. Preparation independence (all state pairs are compatible)
-			* P2+ Def -> 3. Span surjectivity (I.4=H3) (all C is spanned by AxB) 
-			* 3 -> 6. Basis carries over I.7
-			* 1 + 2 -> 4. Injectivity (M total function) I.5=H1
-			* 2 + PI -> 5. Statistical independence I.6 (probability on compatible
-			pairs where one subsystem does not change) 
-			* 5 -> 7. I.8=H2: scalar product carries over (fundamental theorem of
-			projective geometry).
-			* 7 + 6 + 4 (i.e. H1+H2+H3) -> 8. m is tensor product   
-			\end{verbatim}\end{widetext}
-	}
-
-	We start by noting that the tensor product is uniquely
-	characterized, up to isomorphism, by a universal property regarding
-	bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
-	tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
-	map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the property
-	than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ factors
-	through $T$ uniquely.  This means that there exists a {\em unique}
-	$\hat m$, dependent on $m$, such that $\hat m \circ T=m$.  In other
-	words, the following diagram commutes:
-	\begin{center}
-		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{m}"] \\
-			& \mathcal{C}
-		\end{tikzcd}
-	\end{center}
+%%BoundingBox: 45 418 630 588
+%(%old BoundingBox: 45 382 649 547)
+\begin{figure}[ht]
+%\epsfxsize=1.\hsize\leavevmode\epsffile{fig.eps}
+\includegraphics[width=\linewidth, trim={0.2in 5.8in 2.5in 0.1in}, clip=true]{fig.eps}
+\caption{Schematic depiction of the logical implications used
+in this paper. FTPG stands for ``Fundamental Theorem of Projective Geometry''.  \label{f:fig}}\end{figure}
 
 We use the axiomatization of quantum mechanics based on the following
-	postulates (e.g.~\cite{ozawa,nielsenchuang}): (a)~The state of a
+postulates (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The state of a
 system is described by a ray $\pj{\psi}$ corresponding to a set of
@@
 
-	\togli{This axiomatization implicitly contains a definition of ``quantum
-		system'' which is crucial for what follows, so we need to clarify the
-		assumptions that it contains. We will use the following definition for
-		a quantum system\togli{$\stackon[1pt]={\mbox{\tiny
-					def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree of
-			freedom with $d$ (possibly discrete, or continuous, infinite)
+\togli{This axiomatization implicitly contains a definition of
+  ``quantum system'' which is crucial for what follows, so we need to
+  clarify the assumptions that it contains. We will use the following
+  definition for a quantum system\togli{$\stackon[1pt]={\mbox{\tiny
+        def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree
+    of freedom with $d$ (possibly discrete, or continuous, infinite)
     mutually exclusive (commuting) values for each of its properties.
-			Its mathematical description is through a Hilbert space of dimension
-			$d$ which contains all the states that describe the values of its
-			possible properties. In accordance with the postulate (a), these
-			values correspond to a basis of the space, given by the eigenvectors
-			of the observable corresponding to that property''}. \commentg{We may have to revise to be more clear. Where is this used?} }
-	As stated
-	above, we will limit ourselves to kinematically-independent systems,
-	where all state vectors $|\psi\>$ in the system's Hilbert space $\cal H$
-	describe a valid state, {\em unconditioned on anything else}. In
-	particular, we note that restrictions due to superselection rules
-	arise either from practical (not fundamental) limitations on the
-	actions of the experimenter \cite{susskind,zanardi,zanardilloyd} or
-	from the use of ill-defined quantum systems. For example, in
-	situations where indistinguishability plays a role (e.g.~in QFT), one
-	cannot consider an electron as a quantum system: in this case the
-	quantum system is the field. The electron is an excitation of the
-	field, {\em not} a (sub)system.
-
-	This implies that the ``system'' is kinematically independent from
-	anything else.  We call this condition ``preparation independence''.
-	[We emphasize that the kinematic independence is inequivalent to
-	dynamical independence (or isolation).  Indeed if two systems
-	interact, their interaction may lead to dynamical restrictions in the
-	state spaces. We will not consider dynamical evolution in this paper,
-	which is contained in postulate (d).]
+    Its mathematical description is through a Hilbert space of
+    dimension $d$ which contains all the states that describe the
+    values of its possible properties. In accordance with the
+    postulate (a), these values correspond to a basis of the space,
+    given by the eigenvectors of the observable corresponding to that
+    property''}. \commentg{We may have to revise to be more clear.
+    Where is this used?} } As stated above, we will limit ourselves to
+kinematically-independent systems, where all state vectors $|\psi\>$
+in the system's Hilbert space $\cal H$ describe a valid state, {\em
+  unconditioned on anything else}. In particular, we note that
+restrictions due to superselection rules arise either from practical
+(not fundamental) limitations on the actions of the experimenter
+\cite{susskind,zanardi,zanardilloyd} or from the use of ill-defined
+quantum systems. For example, in situations where indistinguishability
+plays a role (e.g.~in QFT), one cannot consider an electron as a
+quantum system: in this case the quantum system is the field. The
+electron is an excitation of the field, {\em not} a (sub)system. We
+call this condition ``preparation independence''.  [We emphasize that
+the kinematic independence is inequivalent to dynamical independence
+(or isolation).  Indeed if two systems interact, their interaction may
+lead to dynamical restrictions in the state spaces. We will not
+consider dynamical evolution in this paper, which is contained in
+postulate (d).]
 
@@
 pairs) and returns a state in the projective space $\pj{\cal C}$ for
-	the composite. The map $M$ that acts on rays corresponds to a map $m:{\cal A}\times{\cal B}\to{\cal C}$ that
-	acts on vectors in the Hilbert spaces $\cal A$, $\cal B$ and $\cal C$.
-	Namely, $\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign
-	indicates the elements in the projective space. We will prove that the
-	map $m$ is the tensor product.
+the composite. The map $M$ that acts on rays corresponds to a map
+$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
+Hilbert spaces $\cal A$, $\cal B$ and $\cal C$.  Namely,
+$\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign indicates the
+elements in the projective space. We will prove that the map $m$ is
+the tensor product. We focus on pure states here: one can easily
+extend to mixed states using standard tools \cite{nielsenchuang}.
 
@@
 $b$ the measurement outcomes of two observables on $\cal A$ and $\cal
-	B$ respectively. The first equality in \eqref{bindip} embodies the
+B$ respectively. The first equality in equation \eqref{bindip} embodies the
 definition of statistical independence, the second follows from the
@@
 \end{align}
-	where the first and second terms contain the scalar product in the composite
+where the first and second terms contain the inner product in the composite
 space $\cal C$. [This is not a new assumption: it follows from the
@@
 either linear or antilinear. This tells us that the corresponding $m$
-	is either linear or antilinear in the first argument. Namely, if
+is either linear or antilinear in the first argument. Namely, if equation
 \eqref{questa} holds, then
@@
 We have obtained the conditions H1, H2 and H3 from the state postulate
-	(a), the measurement postulate (b) and the definition of composite
-	systems. We now prove that these three conditions imply that the (up
-	to now unspecified) composition rule $m$ is the tensor product. More
-	precisely, given a total, span-surjective, bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ that preserves the square of the inner
-	product, we find that $\cal C $ is equivalent to
-	$\cal A \otimes \cal B $ and that $m=\otimes$.
+(a), the measurement postulate (b) and the definitions of composite
+and independent systems. In the methods section we prove that these
+three conditions imply that the (up to now unspecified) composition
+rule $m$ is the tensor product. More precisely, given a total,
+span-surjective, bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$
+that preserves the square of the inner product, we find that $\cal C $
+is equivalent to $\cal A \otimes \cal B $ and that $m=\otimes$.
 
-	Proof. Step 1: the bases of the component systems are mapped to a
-	basis of the composite system. Because of totality property (H1) and
-	because the square of the inner product is preserved, we can conclude
-	that, given two orthonormal bases $\{|a_i\>\}\in{\cal A}$ and
+A few comments on the proof (given below): it is based on the universal
+property of the tensor product, which uniquely characterizes it. First
+we show that the bilinear map $m$ maps subsystems' bases into the
+composite system basis. We also know that there exists a tensor
+product map $ {T}=\otimes$ that can compose the vectors in $\cal A$
+and $\cal B$.  Then we use the universal property: since $m$ is a
+bilinear map, we are assured that there exists a unique $\hat m$ such
+that $\hat m \circ T=m$. Since we show that $\hat m$ is an
+isomorphism, then $\hat{m}$ bijectively maps vectors in $\cal C$ onto
+vectors in the tensor product space. Namely $m={T}=\otimes$.
+
+We conclude with some general comments. The tensor product structure
+of quantum systems is not absolute, but depends on the observables
+that are accessible \cite{zanardi,zanardilloyd}. This is due to the
+fact that an agent that has access to a set of observables will define
+quantum systems differently from an agent that has access to a
+different set of observables. Where one agent sees a single system, an
+agent that has access to less refined observables (and is then limited
+by some superselection rules) can consider the same system as composed
+of multiple subsystems. A typical example \cite{tellerbook} comes from
+quantum field theory. It is customary in basically all quantum optics
+literature to treat different modes of the radiation field (e.g.~the
+output of two lasers) as independent systems composed through the
+tensor product.  Clearly the electromagnetic field is a single system
+and an agent who is able to access an optical mode that is a linear
+combination of the two will give a quantum description for it that
+cannot easily accommodate tensor products. Similarly, an agent can
+consider two electrons as two systems, joined with the tensor product,
+whenever they are distinguishable for all practical purposes (e.g.~the
+electrons are in widely separated physical locations). Yet, in
+principle, electrons are just excitations of a field, and the `true'
+quantum system is the field, not the single electrons
+\cite{teller,tellerbook}.  So, in quantum field theory, the quantum
+systems that should be joined through tensor products are the
+different fields and {\em not} the particles, which are just
+excitations (states) of the fields. In the words of Teller
+(\cite{tellerbook}, pg.22), tensor products can be safely used only if
+there is a ``primitive thisness'', which is captured in the definition
+of system.
+
+It has been pointed out before that the quantum postulates are
+redundant: in \cite{masanes} it was shown that the measurement
+postulate (b) can be derived from the others (a), (c), (d). Here
+instead we have shown how the tensor product postulate (c) can be
+logically derived from the state postulate (a), the measurement
+postulate (b) and a reasonable definition of independent systems, and
+we have described the logical relations among them.  Of course, we do
+not claim that this is the {\em only} way to obtain the tensor product
+postulate from the others.
+
+\vskip 1\baselineskip
+{\par\noindent\bf Methods}
+
+Proof of the statement: {\em given a total, span-surjective, bilinear
+  map $m:{\cal A}\times{\cal B}\to{\cal C}$ that preserves the square
+  of the inner product, we find that $\cal C $ is equivalent to $\cal
+  A \otimes \cal B $ and that $m=\otimes$.}
+
+Step 1: the bases of the component systems are mapped to a basis of
+the composite system. Because of totality property (H1) and because
+the square of the inner product is preserved, we can conclude that,
+given two orthonormal bases $\{|a_i\>\}\in{\cal A}$ and
 $\{|b_j\>\}\in{\cal B}$,
@@
 
-	Step 2: Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is
+Step 2: use the universal property. The tensor product is uniquely
+characterized, up to isomorphism, by a universal property regarding
+bilinear maps: given two vector spaces $\cal A$ and $\cal B$, the
+tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
+map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the property
+than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ factors
+through $T$ uniquely.  This means that there exists a {\em unique}
+$\hat m$, dependent on $m$, such that $\hat m \circ T=m$.  In other
+words, the following diagram commutes:
+\begin{center}
+\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{m}"] \\
+& \mathcal{C}
+\end{tikzcd}
+\end{center}
+Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is
 a bilinear operator (property H2), thanks to the universal property of
@@
 
-	\vskip 1\baselineskip A few comments on the proof: it is based on the
-	universal property of the tensor product, which uniquely characterizes
-	it. In Step 1 we showed that the bilinear map $m$ maps subsystems'
-	bases into the composite system basis. We also know that there exists a
-	tensor product map $ {T}=\otimes$ that can compose the vectors in $\cal A$
-	and $\cal B$. In Step 2 we use the universal property: since $m$ is a
-	bilinear map, we are assured that there exists a unique $\hat m$ such
-	that $\hat m \circ T=m$. Since we show that $\hat m$ is an
-	isomorphism, then $\hat{m}$ bijectively maps vectors in $\cal C$
-	onto vectors in the tensor product space. Namely $m={T}=\otimes$.
-
-	We conclude with some general comments. The tensor product structure
-	of quantum systems is not absolute, but depends on the observables
-	that are accessible \cite{zanardi,zanardilloyd}. This is due to the
-	fact that an agent that has access to a set of observables will define
-	quantum systems differently from an agent that has access to a
-	different set of observables. Where one agent sees a single system, an
-	agent that has access to less refined observables (and is then limited
-	by some superselection rules) can consider the same system as composed
-	of multiple subsystems. A typical example \cite{tellerbook} comes from
-	quantum field theory. It is customary in basically all quantum optics
-	literature to treat different modes of the radiation field (e.g.~the
-	output of two lasers) as independent systems composed through the
-	tensor product.  Clearly the electromagnetic field is a single system
-	and an agent who is able to access an optical mode that is a linear
-	combination of the two will give a quantum description for it that
-	cannot easily accommodate tensor products. Similarly, an agent can
-	consider two electrons as two systems, joined with the tensor product,
-	whenever they are distinguishable for all practical purposes (e.g.~the
-	electrons are in widely separated physical locations). Yet, in
-	principle, electrons are just excitations of a field, and the `true'
-	quantum system is the field, not the single electrons
-	\cite{teller,tellerbook}.  So, in quantum field theory, the quantum
-	systems that should be joined through tensor products are the
-	different fields and {\em not} the particles, which are just
-	excitations (states) of the fields. In the words of Teller
-	(\cite{tellerbook}, pg.22), tensor products can be safely used only if
-	there is a ``primitive thisness'', which is captured in the definition
-	of system.\togli{\comment{forse e' da spiegare meglio}}
-
 \togli{\comment{Commento finale da aggiungere?  E' interessante perche' tutto
@@
 
-	It has been pointed out before that the quantum postulates are
-	redundant: in \cite{masanes} it was shown that the measurement
-	postulate (b) can be derived from the others (a), (c), (d). Here
-	instead we have shown how the tensor product postulate can be
-	logically derived from the state postulate (a), the measurement
-	postulate (b) and a reasonable definition of independent systems, and
-	we have described the logical relations among them.  Of course, we do
-	not claim that this is the {\em only} way to obtain the tensor product
-	postulate from the others.
-
-	L.M. acknowledges useful discussions with M.~Ozawa, P.~Zanardi,
-	S.~Lloyd, D.~Zeh, G.~Auletta, A.~Aldeni. We acknowledge funding from
-	the Attract project through the Eu Horizon 2020 research and
-	innovation programme under grant agreement No 777222.
 \begin{references}
@@
   redundant, Nat. Commun. {\bf 10}, 1361 (2019).
+\bibitem{wootters}W.K. Wootters, Optimal Information Transfer and
+  Real-Vector-Space Quantum Theory. In: Chiribella G., Spekkens R.
+  (eds) Quantum Theory: Informational Foundations and Foils,
+  Fundamental Theories of Physics, vol 181. Springer, Dordrecht
+  (2016).
 \bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum Computation
@@
   quantum mechanics, Am. J. Phys. {\bf 54}, 883 (1986).
-		\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
-		press, 1961).
 \bibitem{peres}A.~Peres, Classical interventions in quantum systems.
@@
   (2015).% https://doi.org/10.1007/JHEP12(2015)099
+\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
 \bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
@@
   {\bf 92},060402 (2004).
+\bibitem{cox}R.T. Cox, The Algebra of Probable Inference (J. Hopkins
+  press, 1961).
+\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
 \bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
@@
     21}, 43 (1991).
-		\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
-		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
 \bibitem{holevo}A. Holevo, Probabilistic and statistical aspects of
@@
 
-	\vskip 5\baselineskip
+\vskip 1\baselineskip
+{\bf Acknowledgements:} L.M. acknowledges useful discussions with
+M.~Ozawa, P.~Zanardi, S.~Lloyd, D.~Zeh, G.~Auletta, A.~Aldeni and
+funding from the Attract project through the Eu Horizon 2020 research
+and innovation programme under grant agreement No 777222. G.C. and
+C.A.A. acknowledge funding from the MCubed program of the University
+of Michigan.
 
-	\section{Appendix: mathematical formulation}\label{app}
-	\setcounter{section}{1}
+* Corresponding Author: maccone@unipv.it.
 
-	The following section needs to distinguish the projective space from
-	the Hilbert space itself. As this is seldom done in quantum mechanics,
-	we review some concepts in that context and introduce the notation
-	that we will be using. If $X$ is a Hilbert space, we denote $\pj{X}$
-	the projective space. The projective space is mathematically
-	constructed from the Hilbert space by removing the origin and
-	quotienting by the equivalence relationship $v \sim \lambda v$, $v\in
-	X$ and $\lambda\in\mathbb{C}$. A quantum state is a point in
-	projective space. Each point of the projective space is called a ray,
-	because for a real vector space to would correspond to a line going through
-	the origin, with the origin removed. As we are in a complex space, the
-	ray should be thought as a complex plane without the origin, which is
-	the space of the vectors reachable from a fixed one through
-	multiplication by a complex number. It can also be thought as a
-	subspace of dimension one.
+\end{document}
 
-	Given a vector $v \in X$, we will denote $\pj{v}$ the ray in the
-	projective space corresponding to $v$. Note that $\pj{v}$ denotes a
-	quantum state, without having picked a modulus or phase. Given two or
-	more vectors $v_1, ..., v_n \in X$, the subspace of $X$ they span
-	(i.e.~all the vectors reached by linear combinations) is noted by
-	$Sp(v_1, ..., v_n)$. Note that this subspace will correspond to a set
-	of rays in the projective space, which we note as $\pj{Sp(v_1, ...,
-		v_n)}$. Given $v,w \in X$, we can write $P(v|w) = \frac{|\< v | w
-		\>|^2}{\< v | v \>\< w | w \>}$ which corresponds to the probability
-	of observing $v$ given $w$ was prepared. Note that $P(v|w) = P(\lambda
-	v| \mu w)$, with non-null $\lambda,\mu\in{\mathbb C}$, and therefore
-	one can write $P(\pj{v} | \pj{w})\equiv P(v|w)$ as a function of the
-	rays.
-
-	\begin{post}\label{post_state}
-		The state of a quantum system is described by a ray $\pj{\psi} = \{
-		\alpha |\psi\> \, |$ non-null $\alpha \in
-		\mathbb{C},|\psi\>\in\mathcal{H}\}$ in a complex Hilbert space
-		$\mathcal{H}$, and the system's observable properties are described
-		by self-adjoint operators acting on that space. %All vectors represent possible   system system states and all self-adjoint operators represent possible system   observables.
-	\end{post}
-
-	\begin{remark}
-		All proofs, except one, should not depend on the dimensionality of
-		the space. The exception is proposition \ref{prop_fundProj} which
-		works if the basis for the space is finite since we are constructing
-		the map one component at a time. It should be possible to extend it
-		using weak convergence in the case of countable basis
-		(i.e. separable Hilbert spaces) by creating a sequence that
-		converges to each vector. We are not specifically addressing this
-		case in this paper.
-	\end{remark}
-
-	\begin{post}\label{post_measurements}
-		The probability that a measurement of a property $X$, described by the operator with
-		spectral decomposition $X = \sum_{x,i }x \frac{| x_i \> \< x_i |}{\< x_i | x_i \>}$  where $i$ is a degeneracy index, returns a value $x$ depends only on $X$ and on the state of the system $\pj{\psi}$ and is given by $P(x|\pj{\psi})=\sum_i \frac{\<\psi| x_i \> \< x_i |\psi\>}{\< \psi | \psi \>\< x_i | x_i \>}$
-		(Born rule).\end{post}
-
-	\begin{defn}[Compatible states]\label{def_compatible}
-		Let A and B be two systems. Let ${\mathcal{A}}$ and ${\mathcal{B}}$
-		be their corresponding state spaces. We say two (pure) states $(\pj{a},
-		\pj{b}) \in \pj{\mathcal{A}} \times \pj{\mathcal{B}}$ are compatible
-		iff the respective systems can be prepared in such states at the
-		same time. Formally, the proposition $\pj{a} \wedge \pj{b}$ is
-		possible, which means it does not correspond to the empty set in the
-		$\sigma$-algebra of the probability space\footnote{The impossible event is not an event with probability zero, rather it is an event that cannot be created at all.
-For example, ``the dice shows a number that is even and less than two'', or ``the electron is prepared in spin up along $x$ and also along $z$'' are impossible events.}.
-	\end{defn}
-
-	{Note that the proposition $\pj{a} \wedge \pj{b}$ refers to either
-		preparation or measurement of both systems  in the respective state.
-		The proposition $\pj{a} | \pj{b}$, instead, is the one that
-		corresponds to preparing one system in one state and measuring the
-		other system in the other state \cite{cox}. Incompatible states
-		refer to non-commuting observables.}
-
-	\begin{defn}[Preparation independence]\label{def_indep}
-		Two systems are said independent iff the preparation of one does not affect the preparation of the other.
-Formally, all (pure) state pairs $(\pj{a}, \pj{b}) \in \pj{\mathcal{A}}\times \pj{\mathcal{B}}$ are compatible.
-	\end{defn}
-
-	\begin{prop}\label{prop_singleBorn}
-		Given two systems, each prepared independently in their own state, the probability of measuring a value for one system depends only on the preparation of that system.
-That is, $P(\pj{a_1}|\pj{a_2}\wedge \pj{b})=P(\pj{a_1}|\pj{a_2})$.
-	\end{prop}
-	\begin{proof}
-		We first note that, by postulate \ref{post_measurements}, the probability of measuring a value for one system depends only on the preparation of that system, which means that it is independent of the properties of any other system.
-Therefore $P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = \frac{\<a_1| a_2 \> \< a_2 |a_1\>}{\< a_1 | a_1 \>\< a_2 | a_2 \>} = P(\pj{a_1} | \pj{a_2})$
-	\end{proof}
-
-	\begin{defn}[Composite systems]\label{def_comp}
-		Let A and B be two systems. The composite system C of A and B is formed by the simple collection of those and only those two systems, in the sense that it satisfies the following two requirements.
-		\begin{enumerate}
-			\item Every preparation of both subsystems is a preparation of the
-			composite. Formally, let $\pj{\mathcal{C}}$ be the state space
-			for C, there exists a map (not yet specified)
-			$M:\pj{\mathcal{A}}\times\pj{\mathcal{B}}\to\pj{\mathcal{C}}$
-			such that, for any compatible pair of (pure) states $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$, the proposition $\pj{a} \wedge \pj{b}$ is equivalent to the (pure) state $M(\pj{a},\pj{b}) \in \pj{\mathcal{C}}$ where $M$ returns the state of the composite system where the subsystems were prepared in the given states.
-In other words, $\pj{a} \wedge \pj{b}$ and $M(\pj{a},\pj{b})$ correspond to the same event in probability space\footnote{We will end up proving that the map $M$ leads to the tensor product.}.
-			\item Every preparation of the composite gives a non-trivial measurement on the components.
-Formally, for every $\pj{c} \in \pj{\mathcal{C}}$, we can find at least $\pj{a} \in \pj{\mathcal{A}}$ and $\pj{b} \in \pj{\mathcal{B}}$ such that $P(\pj{a} \wedge \pj{b}|\pj{c})\neq 0$.
-		\end{enumerate}
-		Requirement 1 ensures that the composite system contains all the properties of the components.
-Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components.
-	\end{defn}
-
-	\begin{prop}[Span surjectivity, H3]\label{prop_spanSurj}
-		The map $M :
-		\pj{\mathcal{A}}\times\pj{\mathcal{B}} \to \pj{\mathcal{C}}$ is span
-		surjective, meaning that the span of the image coincides with the
-		whole space. That is $Sp(\{ c \in \mathcal{C} \, | \, \pj{c} \in
-		M(\pj{\mathcal{A}}, \pj{\mathcal{B}})\}) = \mathcal{C}$.
-	\end{prop}
-	\begin{proof}
-		Consider $I=\{ c \in \mathcal{C} \, | \, \pj{c} \in M(\pj{\mathcal{A}}, \pj{\mathcal{B}})\}$ and its span.
-This forms a subspace of $\mathcal{C}$.
-By requirement 2 of \ref{def_comp}, for any $c \in \mathcal{C}$ we can always find $a \in \mathcal{A}$ and $b \in \mathcal{B}$ such that $P(\pj{a} \wedge \pj{b} | \pj{c} )\neq 0$.
-This means there is no element in $\mathcal{C}$ that is orthogonal to $Sp(I)$, therefore $Sp(I)$ must cover the whole $\mathcal{C}$.
-	\end{proof}
-
-	\begin{prop}[Totality, H1]\label{prop_totality}
-		The map $M$ is in general a partial function.\footnote{A partial
-			function is one that is not defined on the full domain. For
-			example, $\sqrt(x)$ is a partial function since is not defined for
-			$x<0$.} However, if A and B are independent, $M$ is a total function.\footnote{A total function is one that is defined on the full domain.
-For example, $x^2$ is a total function since it is defined for any $x$.}
-	\end{prop}
-	\begin{proof}
-		As $M(\pj{a},\pj{b})$ is defined only if $(\pj{a},\pj{b}) \in \pj{\mathcal{A}}\times\pj{\mathcal{B}}$ are a compatible pair of pure states, it is not defined on pairs that are not compatible.
-If the two systems are independent, however, all pairs are allowed and $M$ is a total function.
-	\end{proof}
-
-	\begin{remark}
-		As noted in \ref{def_comp}, if $\pj{a}$ and $\pj{b}$ are incompatible, $\pj{a} \wedge \pj{b}=\emptyset$ corresponds to the impossible event (i.e.~the empty set in the $\sigma$-algebra).
-This is not a state, and therefore $M(\pj{a},\pj{b})$ is not defined on incompatible pairs.
-
-		However, in the end we will construct a map $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ on the vector spaces.
-There the zero vector plays the role of the impossible event.
-Therefore independent systems will map each pair to a non-zero element of the tensor product, while systems that are not independent will map incompatible states to the zero vector (e.g.
-the composite state of two electrons will exclude the cases where both electrons are in the same state).
-	\end{remark}
-
-	\begin{prop}[Statistical independence]\label{prop_statInd}
-		Let $\pj{\mathcal{A}}$ and $\pj{\mathcal{B}}$ be the state spaces of two quantum systems and $\pj{\mathcal{C}}$ be the state space of their composite.
-The map $M : \pj{\mathcal{A}} \times \pj{\mathcal{B}} \to \pj{\mathcal{C}}$ is such that:
-		\begin{align}
-		P(M(\pj{a_1},\pj{b}) | M(\pj{a_2},\pj{b})) = P(\pj{a_1} | \pj{a_2})  \\ P(M(\pj{a},\pj{b_1}) | M(\pj{a},\pj{b_2})) = P(\pj{b_1} | \pj{b_2})
-		\end{align}
-		for all $a, a_1, a_2 \in \mathcal{A}$ and $b, b_1, b_2 \in \mathcal{B}$
-	\end{prop}
-	\begin{proof}
-		By \ref{prop_singleBorn} we have $P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = P(\pj{a_1} | \pj{a_2})$ and similarly $P(\pj{b_1} | \pj{a} \wedge \pj{b_2}) = P(\pj{b_1} | \pj{b_2})$.
-Using standard probability rules and remembering that $M(\pj{a},\pj{b})\equiv \pj{a} \wedge \pj{b}$ by \ref{def_comp}, we have $P(M(\pj{a_1},\pj{b}) | M(\pj{a_2},\pj{b})) = P(\pj{a_1} \wedge \pj{b} | \pj{a_2} \wedge \pj{b})  = P(\pj{b} | \pj{a_2} \wedge \pj{b}) P(\pj{a_1} | \pj{a_2} \wedge \pj{b} \wedge \pj{b})  = P(\pj{b} | \pj{b}) P(\pj{a_1} | \pj{a_2} \wedge \pj{b}) = P(\pj{a_1} | \pj{a_2})$, since trivially $P(\pj{b}|\pj{b})=1$.
-Similarly $P(M(\pj{a},\pj{b_1}) | M(\pj{a},\pj{b_2})) =P(\pj{a} \wedge \pj{b_1} | \pj{a} \wedge \pj{b_2}) = P(\pj{b_1} | \pj{b_2})$
-	\end{proof}
-
-	\begin{prop}[Fundamental theorem of projective geometry]\label{prop_fundProj}
-		Let $X$ and $Y$ be two complex Hilbert spaces and $\pj{X}$ and $\pj{Y}$ their respective projective spaces.
-Let $M : \pj{X} \to \pj{Y}$ be a map such that $P(\pj{v}|\pj{w}) = P(M(\pj{v})|M(\pj{w}))$.
-Then we can find, up to a total phase, a unique map $m : X \to Y$ such that $\pj{m(v)}=M(\pj{v})$.
-Moreover, $m$ is either linear, $\<v|w\> = \<m(v)|m(w)\>$, or anti-linear, $\<v|w\> = \<m(w)|m(v)\>$.
+\begin{prop}[Mixed states]
+  Up to now we have only considered pure states, described by rays
+  $\pj{\psi}$ or vectors $|\psi\>$ in the Hilbert space.  Using
+  standard techniques we can extend the analysis to more general mixed
+  states, described by a density matrix $\rho$. 
 \end{prop}
 
-	\begin{remark}
-		The above proposition is, for the most part, an adaptation of the fundamental theorem of projective geometry \cite{fun}.
-The conservation of the probability imposes the semi-linear map to be either linear or anti-linear (i.e.
-conjugate-linear).
-	\end{remark}
+Using the Born rule (which, we remind, is one of the postulates we
+employ), it is simple to show that a density matrix (i.e.~a positive,
+trace one operator) can be given the interpretation of the system
+being in a pure state with some probability. Indeed, any density
+matrix can be always written as $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$,
+where the $|\psi_i\>$ are not necessarily orthogonal and where
+$p_i\geqslant 0$ sum up to one: $\sum_ip_i=1$. We now show that, a
+system which is in a state $|\psi_i\>$ with probability $p_i$ can be
+described by a density matrix $\rho$ \cite{nielsenchuang}.
 
-	\begin{proof}
-		First we note that, given an orthonormal basis $\{e_i\}_{i \in I}$ over $X$, we can use $M$ to construct a corresponding basis over $Y' \subseteq Y$ where $\pj{Y'} = M(\pj{X})$.
-In fact, for each $\pj{e_i}$, pick a unit $u_i \in M(\pj{e_i})$.
-We have $\delta_{ij} = |\<e_i | e_j \>|^2 = P(e_i | e_j) = P(\pj{e_i}|\pj{e_j}) = P(M(\pj{e_i})|M(\pj{e_j})) = P(\pj{u_i}|\pj{u_j})= |\<u_i | u_j \>|^2$.
-The set $\{u_i\}_{i \in I}$ spans the entire $Y'$ since for all $y \in Y'$ we can find $x \in X$ and at least one $u_i$ such that $|\<y | u_i \>|^2 = P(\pj{y}|\pj{u_i}) = P(M(\pj{x})|M(\pj{e_i})) = P(x | e_i) = |\<x | e_i \>|^2 \neq 0$.
-Note that we have an arbitrary choice for each $u_i$, since we have to pick a vector from the unit circle (i.e.
-a phase for each basis vector).
-This corresponds to a choice of gauge.
+Consider an arbitrary observable with eigenstate $|a\>$. The Born rule
+says that, the probability of finding the related eigenvalue $a$
+given that the system is in a state described by a normalized vector
+$|\psi_i\>$, is $p(a|i)=|\<\psi_i|a\>|^2$. But if the system is in
+state $|\psi_i\>$ with probability $p_i$, a trivial application of the
+Bayes rule, marginalizing over $i$, tells us that
+\begin{align} &p(a)=\sum_ip(a\wedge
+  i)=\sum_ip(a|i)p_i=\sum_ip_i|\<\psi_i|a\>|^2
+\nonumber\\&=
+  \mbox{Tr}[|a\>\<a|\sum_ip_i|\psi_i\>\<\psi_i|]=\mbox{Tr}[|a\>\<a|\rho]
+\labell{bo}\;.
+\end{align}
+As is customary, we interpret the last term as the Born rule for
+density matrices, which then says that a system described by the
+density matrix $\rho$ is equivalent (thanks to the arbitrariness of
+$|a\>$) to a system described by a state $|\psi_i\>$ with probability
+$p_i$. Note that the decomposition $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$
+of a mixed state $\rho$ into pure states $|\psi_i\>,p_i$ is not
+unique, except for the pure states themselves where one of the $p_i$
+is one.  This is an interesting peculiar characteristic of quantum
+mechanics.
 
-		We also note that the map is colinear, meaning that if $U_X, V_X \subseteq X$ are two subgroups such that $U_X \subset V_X$, then $U_Y, V_Y \subseteq Y$ such that $\pj{U_Y} = M(\pj{U_X})$ and $\pj{V_Y} = M(\pj{V_X})$ are subgroups of $Y$ and $U_Y \subset V_Y$.
-In fact, take a basis $\{e_i\}_{i \in I} \subset X$ such that $\{e_i\}_{i \in I_U \subset I} \subset \{e_i\}_{i \in I_V \subset I}$ are bases for $U_X$ and $V_X$ respectively.
-An element of $X$ belongs to $U_X$ if and only if it is not orthogonal only to elements of the basis of $U_X$ and belongs to $V_X$ only if it not orthogonal only to elements of the basis of $V_X$.
-As the map $M$ preserves orthogonality, these relationships are preserved by the map.
-Therefore $U_Y$ and $V_Y$ are subgroups of $Y$ such that $U_Y \subset V_Y$.
+In the discussion above, we have not specified the Hilbert space where
+the vectors live. If we suppose that $|\psi_i\>$ are vectors in the
+tensor product Hilbert space of two subsystems, then the discussion we
+have given for pure states in our paper trivially extends to mixed
+states: namely we can conclude that the mixed states of composite
+systems are density matrices (positive operators with trace one) that
+act on the tensor product Hilbert space.
 
-		We now use the gauge freedom to redefine the basis such that for all $i$ we have $M(\pj{e_i}) = \pj{v_i}$ and $M(\pj{e_1 + e_i}) = \pj{v_1 + v_i}$.
-Let $v_1 = u_1$.
-This is the only arbitrary choice we make, and corresponds to the choice of a global phase.
-For each $i>1$, consider $e_1 + e_i$.
-This will belong to the subspace $Sp(e_1, e_i)$.
-This subspace, when mapped through $M$, will give us the subspace spanned by $v_1$ and $u_i$.
-That is, $M(\pj{Sp(e_1, e_i)})=\pj{Sp(v_1, u_i)}$.
-This means we can find a unique $k \in \mathbb{C}$ such that $M(\pj{e_1+e_i})=\pj{v_1+ku_i}$.
-We fix $v_i = k u_i$.
-Note that $P(\pj{e_1} | \pj{e_1+e_i}) = \frac{1}{2}=  P(\pj{e_i} | \pj{e_1+e_i})=P(\pj{v_1} | \pj{v_1+k u_i})=P(\pj{u_i} | \pj{v_1+k u_i})$.
-Therefore $|k| = 1$ and $k u_i = v_i$ is a unit vector.
-
-		Now we want to show that $M(\pj{e_1 + c e_i})=\pj{v_1 + \tau_i(c)
-			v_i}$ where either $\tau_i(c) = c$ or $\tau_i(c) = c^\dagger$. For
-		each $i$, consider $w = e_1 + c e_i \in Sp(e_1, e_i)$. Since
-		$M(\pj{w}) \subset \pj{Sp(v_1, v_i)}$, there must be a $\tau_i(c)$
-		such that $\pj{v_1 + \tau_i(c) v_i} = M(\pj{w})$. Since we must have
-		$P(\pj{e_i} | \pj{w}) = P(\pj{v_i} | M(\pj{w}))$ and $P(\pj{e_1 +
-			e_i} | \pj{w}) = P(\pj{v_1 + v_i} | M(\pj{w}))$, we must have
-		$|c| = |\tau_i(c)|$ and $cos(arg(c)) = cos(arg(\tau_i(c)))$ for any
-		$c$. This means that either $\tau_i(c) = c$ or $\tau_i(c) =
-		c^*$.
-
-		Next we want to show that $\tau_i(c) = \tau_j(c)$ for all pairs $(i,j)$.
-That is, either we have to take the complex conjugate of all components or of none.
-Consider $e_i - e_j$.
-We have $\pj{e_i - e_j} \subset \pj{Sp(e_i, e_j)}$ and, for any $c \in \mathbb{C}$, $\pj{e_i - e_j} \subset \pj{Sp(e_1 + c e_i, e_1 + c e_j)}$.
-By construction, we have $M(\pj{e_i - e_j})\subset \pj{Sp(v_i, v_j)}$ and $M(\pj{e_i - e_j})\subset \pj{Sp(v_1 + \tau_i(c) v_i, v_1 + \tau_j(c) v_j)}$.
-Therefore $M(\pj{e_i - e_j}) = \pj{Sp(v_i, v_j)} \cap \pj{Sp(v_1 + \tau_i(c) v_i, v_1 + \tau_j(c) v_j)} = \pj{\tau_i(c) v_i - \tau_j(c) v_j}$.
-This means that, for all $c$, $\tau_i(c) = \tau_j(c)$.
-
-		Now we show that for all $c_2, ..., c_n \in \mathbb{C}$ we have $M(\pj{e_1 + c_2 e_2 + ...
-+ c_n e_n}) = \pj{v_1 + \tau(c_2) v_2 + ...
-+ \tau(c_n) v_n}$.
-We prove this by induction.
-If only the first two components are non-zero, we have  $M(\pj{e_1 + c_2 e_2}) = \pj{v_1 + \tau(c_2) v_2}$ by construction.
-Let $2 < p \leq n$.
-If we assume $M(\pj{e_1 + c_2 e_2 + ...
-+ c_{p-1} e_{p-1}}) = \pj{v_1 + \tau(c_2) v_2 + ...
-+ \tau(c_{p-1}) v_{p-1}}$, then $M(\pj{e_1 + c_2 e_2 + ...
-+ c_p e_p}) \subset M(\pj{Sp(e_1 + c_2 e_2 + ...
-+ c_{p-1} e_{p-1}, e_p})) = \pj{Sp(v_1 + c_2 v_2 + ...
-+ c_{p-1} v_{p-1}, v_p)}$.
-This means that there exists $k_p \in \mathbb{C}$ such that $M(\pj{e_1 + c_2 e_2 + ...
-+ c_p e_p}) = \pj{v_1 + c_2 v_2 + ...
-+ c_{p-1} v_{p-1} + k_p v_p)}$.
-But we also have $M(\pj{e_1 + c_2 e_2 + ...
-+ c_p e_p}) \subset M(\pj{Sp(e_1 + c_p e_p, c_2 e_2 + ...
-+ c_{p-1} e_{p-1}})) = \pj{Sp(v_1 + \tau(c_p) v_p, c_2 v_2 + ...
-+ c_{p-1} v_{p-1}, v_p)}$.
-The only way this can work is if $k_p = \tau(c_p)$.
-
-		We also need to show the above works when there is no component on the first element of the basis.
-That is, for all $c_2, ..., c_n \in \mathbb{C}$ we have $M(\underbar{$c_2 e_2 + ...
-+ c_n e_n$}) = \underbar{$\tau(c_2) v_2 + ...
-+ \tau(c_n) v_n$}$.
-First note that $M(\underbar{$c_2 e_2 + ...
-+ c_n e_n$}) \subset M(\underbar{$Sp(e_2, ..., e_n)$)} = \underbar{$Sp(v_2, ..., v_n)$}$.
-Also note that $M(\underbar{$c_2 e_2 + ...
-+ c_n e_n$}) \subset M(\underbar{$Sp(e_1, e_1 + c_2 e_2 + ...
-+ c_n e_n)$)} = \underbar{$Sp(v_1, v_1 + \tau(c_2) v_2 + ...
-+ \tau(c_n) v_n)$}$.
-The only way this can work is if $M(\underbar{$c_2 e_2 + ...
-+ c_n e_n$}) = \underbar{$\tau(c_2) v_2 + ...
-+ \tau(c_n) v_n$}$.
-
-		We can now define $m : X \to Y$ such that $m(e_i) = v_i$ for all $i$ and $m(\sum_{i \in I} c_i e_i) = \sum_{i \in I} \tau(c_i) v_i$.
-This means $\underbar{$m(\sum_{i \in I} c_i e_i)$} = M(\underbar{$\sum_{i \in I} c_i e_i$})$.
-Moreover, if $\tau(c) = c$ we have $\<m(\sum_{i \in I} c_i e_i)|m(\sum_{j \in I} d_j e_j)\> = \<\sum_{i \in I} c_i v_i|\sum_{j \in I} d_j v_j\> = c_i^* d_j \delta_{ij} = \<\sum_{i \in I} c_i e_i|\sum_{j \in I} d_j e_j\>$.
-On the other hand, if $\tau(c) = c^*$ we have $\<m(\sum_{i \in I} c_i e_i)|m(\sum_{j \in I} d_j e_j)\> = \<\sum_{i \in I} c_i^* v_i|\sum_{j \in I} d_j^* v_j\> = \<\sum_{j \in I} d_j v_j|\sum_{i \in I} c_i v_i\> = d_i^* c_j \delta_{ij} = \<\sum_{j \in I} d_j e_j|\sum_{i \in I} c_i e_i\>$.
-	\end{proof}
-
-	\begin{remark}
-		The fact that the proposition identifies either a linear map or an anti-linear (i.e.~conjugate-linear)  has a clear physical interpretation.
-Since, for a Hilbert space, the conjugate vector space is equivalent to the dual space, a conjugate-linear map is one that preserves the inner product but maps ket vectors into bra vectors.
-Looking ahead, the above result does not exclude a composition map similar to the tensor product, but that maps the kets of one or both subsystems into bras in the composite system.
-
-		Such a conjugate-linear map is physically indistinguishable from a map that maps kets into kets: the physics is only contained in the probabilities which are the square moduli of inner products.
-The idea is that, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear.
-With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.
-
-		Note that this unnecessary subtlety could in principle be avoided by
-		reformulating quantum mechanics in terms of quantum states given by
-		density matrices $\rho=|\psi\>\<\psi|$ (which contain both  kets and
-		bras), as is done, for example in \cite{ozawa,holevo}. In this paper
-		we employed the more familiar formulation in which quantum states are
-		rays in Hilbert space (identified either by kets or bras).
-	\end{remark}
-
-	\begin{prop}[Bilinearity, H2]\label{prop_bilinearity}
-		Let $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ be a map such that, for all $(a,b) \in \mathcal{A} \times \mathcal{B}$, we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$.
-Then $m$ must be bilinear.
-That is:
-		\begin{align}
-		m(k_1a_1 + k_2a_2, b)=k_1m(a_1, b) + k_2m(a_2, b) \\
-		m(a, k_1b_1 + k_2b_2)=k_1m(a, b_1) + k_2m(a, b_2)
-		\end{align}
-		for all $a, a_1, a_2 \in \mathcal{A}$, $b, b_1, b_2 \in \mathcal{B}$ and $k_1, k_2 \in \mathbb{C}$.
-	\end{prop}
-
-	\begin{proof}
-		If we fix $b \in \mathcal{B}$, then we have $M_b : \pj{\mathcal{A}} \to \pj{\mathcal{C}}$ where $M_b(\pj{a}) = M(\pj{a}, \pj{b})$.
-By \ref{prop_statInd} and \ref{prop_fundProj} we can find a linear map $m_b : \mathcal{A} \to \mathcal{C}$ such that $\pj{m_b(a)} = M_b(\pj{a}) = M(\pj{a}, \pj{b})$.
-As this must map subspace to subspace, we must have $m(a, b) = k m_b(a)$ for some $k \in \mathbb{C}$.
-Since $m_b$ is linear, we have $m(k_1a_1 + k_2a_2, b)=k_1m(a_1, b) + k_2m(a_2, b)$ for any $a_1, a_2 \in \mathcal{A}$ and $k_1, k_2 \in \mathbb{C}$.
-We can repeat the argument fixing $a \in \mathcal{A}$, and find $m(a, k_1b_1 + k_2b_2)=k_1m(a, b_1) + k_2m(a, b_2)$ for any $b_1, b_2 \in \mathcal{B}$ and $k_1, k_2 \in \mathbb{C}$.
-	\end{proof}
-
-	\begin{prop}[Subsystems' basis gives composite system
-		basis]\label{prop_basis}
-		Let $\{a_i\}_{i\in I}$ and $\{b_j\}_{j \in J}$ be bases of $\mathcal{A}$ and $\mathcal{B}$ respectively, then a set of unit vectors $\{e_{ij}\}_{(i,j) \in I \times J} \subset \mathcal{C}$  such that $e_{ij} \in M(\underbar{$a_i$}, \underbar{$b_j$})$ forms a basis for $\mathcal{C}$.
-	\end{prop}
-
-	\begin{proof}
-
-		Since $M$ is a map on the projective spaces, it maps spans to spans.
-		Since the span of the basis of $\mathcal{A}$ and $\mathcal{B}$ is
-		the whole space, then the span of the image of the basis is the
-		whole image of $M$. By \ref{prop_spanSurj}, the image of $M$
-		coincides with the whole $\mathcal{C}$. Therefore, given
-		$\{a_i\}_{i\in I}$ and $\{b_j\}_{j \in J}$ bases of $\mathcal{A}$
-		and $\mathcal{B}$ respectively, any set of unit vectors
-		$\{e_{ij}\}_{(i,j) \in I \times J} \subset \mathcal{C}$ such that
-		$e_{ij} \in M(\pj{a_i}, \pj{b_j})$ spans the whole $\mathcal{C}$.
-
-		Now consider $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))$.
-If $i=k$ and $j=l$ we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))= P(M(\pj{a_i},\pj{b_j})|M(\pj{a_i},\pj{b_j})) = 1$.
-If $i\neq k$, we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l})) = P(\pj{a_i} \wedge \pj{b_j} | \pj{a_k} \wedge \pj{b_l}) \leq P(\pj{a_i} | \pj{a_k} \wedge \pj{b_l})$.
-By \ref{prop_singleBorn} we have $P(\pj{a_i} | \pj{a_k} \wedge \pj{b_l}) = P(\pj{a_i} | \pj{a_k}) = 0$ since $a_i$ and $a_k$ are different elements of an orthogonal basis.
-Therefore we have $P(M(\pj{a_i},\pj{b_j})|M(\pj{a_k},\pj{b_l}))=\delta_{ik}\delta_{jl}$ which means $\<e_{ij} | e_{kl} \> = \delta_{ik}\delta_{jl}$.
-
-		The elements $e_{ij}$ form a set of orthonormal vectors that span the whole space and are therefore a basis.
-	\end{proof}
-
-	\begin{thrm}[Composite system theorem]\label{theo}
-		The state space of a composite system of independent systems is given by the tensor product of the spaces of the component systems.
-	\end{thrm}
-
-	\begin{proof}
-		We are looking for a map $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ such that, for all $(a,b) \in \mathcal{A} \times \mathcal{B}$ we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$.
-We saw in \ref{prop_bilinearity} that if $m$ exists, it must be bilinear.
-
-		Now we show that, if $m$ exists, then $\mathcal{C} \cong \mathcal{A}
-		\otimes \mathcal{B}$ (where $\cong$ indicates an isomorphism) and $m :
-		\mathcal{A} \times \mathcal{B} \to \mathcal{A} \otimes \mathcal{B}$ is
-		the standard map from the scalar product to the tensor product. As $m
-		: \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a bilinear
-		operator, by the universal property of the tensor product we can find
-		a linear operator $\hat{m} : \mathcal{A} \otimes \mathcal{B} \to
-		\mathcal{C}$ such that $m(a, b) = \hat{m}(a \otimes b)$. By
-		\ref{prop_basis} the set $\{ m(a_i, b_j)\}_{(i,j) \in I \times J}$
-		forms a basis since $ m(a_i, b_j) \in M(\pj{a_i}, \pj{b_j})$ for all
-		$(i, j)$, therefore $\hat{m}(\{ a_i \otimes b_j\}_{(i,j) \in I \times
-			J})$ also forms a basis since $\hat{m}(a_i\otimes b_j)=m(a_i, b_j)$.
-		By \ref{prop_totality}, each $m(a_i,b_j)$ will correspond to a unit
-		vector in $\mathcal{C}$. We have $\<\hat m(a_i\otimes b_j)| \hat
-		m(a_k\otimes b_\ell)\>_\mathcal{C} =\<m(a_i, b_j)| m(a_k,
-		b_\ell)\>_\mathcal{C} = \delta_{ik}\delta_{j\ell} = \<a_i\otimes b_j|
-		a_k \otimes b_\ell\>_{\otimes}$. The function $\hat{m}$, then,
-		preserves the inner product across all elements of the basis and is
-		therefore an isomorphism for Hilbert spaces. We have $\mathcal{C}
-		\cong \mathcal{A} \otimes \mathcal{B}$ and $m(a, b) = \hat{m}(a
-		\otimes b) \cong a \otimes b$.
-
-		Given that the tensor product map exists and it satisfies all the properties $m$ must satisfy, then $m$ exists and it is the tensor product.
-	\end{proof}
-
-	To conclude, we give a schematic outline of the logical implications
-	that led us to the result. This is an expanded version of Fig.~1 of
-	the main paper:
-	\begin{enumerate}
-		\item P\ref{post_state}: states and observables postulate.
-
-		\item P\ref{post_measurements}: Born rule (measurement postulate).
-
-		\item Def \ref{def_indep}: Preparation independence: systems are independent if the preparation of one does not affect the other.
-
-		\item P\ref{post_measurements} $\Rightarrow$ \ref{prop_singleBorn}: the outcome probabilities depend only on the
-		inner product.
-
-		\item Def \ref{def_comp}: Composite system definition: A composite system is a
-		collection of the subsystems (i.e.~all compatible states give a preparation) and only of the subsystems
-		(i.e.~all composite preparations give non-trivial measurements on the
-		subsystems).
-
-		\item P\ref{post_measurements} + Def \ref{def_comp} $\Rightarrow$ \ref{prop_spanSurj} (H3): Span
-		surjectivity (all composiste $\cal C$ are superpositions of ${\cal A}$ and ${\cal B}$).
-
-		\item Def \ref{def_indep} + Def \ref{def_comp} $\Rightarrow$ \ref{prop_totality} (H1):
-		Totality (all possible state pairs of the subsystems correspond to a state of the composite).
-
-		\item P\ref{post_state} + \ref{prop_singleBorn} +  Def \ref{def_comp} $\Rightarrow$ \ref{prop_statInd}: Statistical
-		independence (if one subsystem does not change, the probability on the composite system is given by the probability of the subsystem that changes).
-
-		\item \ref{prop_fundProj}: Fundamental theorem of projective geometry (preserving square of inner product leads to unique linear map)
-
-		\item \ref{prop_statInd} + \ref{prop_fundProj} $\Rightarrow$ \ref{prop_bilinearity} (H2) composition map on vector spaces is bilinear.
-
-		\item P\ref{post_measurements} + Def \ref{def_comp} + \ref{prop_spanSurj} (H3) $\Rightarrow$ \ref{prop_basis}: Basis
-		carries over from subsystems to composite
-
-		\item \ref{prop_totality} + \ref{prop_bilinearity} + \ref{prop_basis} $\Rightarrow$ \ref{theo}: the composition map is the tensor product.
-	\end{enumerate}
-\end{document}
+\vskip 1\baselineskip

### 2020-03-31 Articles/2019-QuantumComposition/Submissions/NatureComm/supplementaryinfo.tex
Extending the proof to countable case

@@
 		\alpha |\psi\> \, |$ non-null $\alpha \in
-		\mathbb{C},|\psi\>\in\mathcal{H}\}$ in a complex Hilbert space
+		\mathbb{C},|\psi\>\in\mathcal{H}\}$ in a separable complex Hilbert space
 		$\mathcal{H}$, and the system's observable properties are described
@@
 	\begin{remark}
-		All proofs, except one, should not depend on the dimensionality of
-		the space. The exception is proposition \ref{prop_fundProj} which
-		works if the basis for the space is finite since we are constructing
-		the map one component at a time. It should be possible to extend it
-		using weak convergence in the case of countable basis
-		(i.e. separable Hilbert spaces) by creating a sequence that
-		converges to each vector. We are not specifically addressing this
-		case in this paper.
+		All proofs, except one, do not depend on the dimensionality of
+		the space.
+The exception is proposition \ref{prop_fundProj} for which the prove the finite case by induction and then show that it holds in the limit.
+This would not work in the non-separable case, since the basis would not be countable.
 	\end{remark}
@@
 	\begin{prop}[Fundamental theorem of projective geometry]\label{prop_fundProj}
-		Let $X$ and $Y$ be two complex Hilbert spaces and $\pj{X}$ and $\pj{Y}$ their respective projective spaces.
+		Let $X$ and $Y$ be two separable complex Hilbert spaces and $\pj{X}$ and $\pj{Y}$ their respective projective spaces.
 Let $M : \pj{X} \to \pj{Y}$ be a map such that $P(\pj{v}|\pj{w}) = P(M(\pj{v})|M(\pj{w}))$.
@@
 
+		The above works also over a countable sum.
+That is, for all $c_2, ..., c_n, ...
+\in \mathbb{C}$ we have $M(\pj{e_1 + c_2 e_2 + ...
++ c_n e_n + ...}) = \pj{v_1 + \tau(c_2) v_2 + ...
++ \tau(c_n) v_n + ...}$.
+Let $X$ be a separable space.
+Let $a = \sum_{k=1}^\infty c_k e_k$ such that $c_1 = 1$.
+Let $a_i = \sum_{k=1}^i c_k e_k$ be the sum of the first $i$ components.
+We have $\lim\limits_{i \to \infty} a_i = a$.
+Let $b_i = \sum_{k=1}^i \tau(c_k) v_k$ and $b = \sum_{k=1}^\infty \tau(c_k) e_k$.
+We have $\lim\limits_{i \to \infty} b_i = b$.
+We already know that $\pj{b_i} = M(\pj{a_i})$ for all finite $i$.
+We need to show that $\pj{b} = M(\pj{a})$.
+First note that, given $a,b \in Y$, $a=b$ if and only if $\<a,c\> = \<b,c\>$ for all $c \in Y$.
+Therefore $\pj{a}=\pj{b}$ if and only if $P(\pj{a}, \pj{c}) = P(\pj{b}, \pj{c})$ for all $c \in Y$.
+For all $c \in X$ we have $\lim\limits_{i \to \infty} P(\pj{a_i}, \pj{c}) = P(\pj{a}, \pj{c}) = P(M(\pj{a}), M(\pj{c}))$ we also have  $\lim\limits_{i \to \infty} P(\pj{a_i}, \pj{c}) = \lim\limits_{i \to \infty} P(M(\pj{a_i}), M(\pj{c})) = \lim\limits_{i \to \infty} P(b_i, M(\pj{c})) = P(\pj{b}, M(\pj{c}))$.
+Therefore $P(M(\pj{a}), M(\pj{c})) = P(\pj{b}, M(\pj{c}))$ for all $c \in X$.
+Note that $M$ is bijective over $Y'=M(\pj{X})$.
+Therefore $P(M(\pj{a}), \pj{c}) = P(\pj{b}, \pj{c})$ for all $c \in Y$ and $\pj{b} = M(\pj{a})$.
+
 		We also need to show the above works when there is no component on the first element of the basis.
@@
 + \tau(c_n) v_n$}$.
+With same reasoning as before, we can extend the sum to the countably infinite case.
 
@@
 On the other hand, if $\tau(c) = c^*$ we have $\<m(\sum_{i \in I} c_i e_i)|m(\sum_{j \in I} d_j e_j)\> = \<\sum_{i \in I} c_i^* v_i|\sum_{j \in I} d_j^* v_j\> = \<\sum_{j \in I} d_j v_j|\sum_{i \in I} c_i v_i\> = d_i^* c_j \delta_{ij} = \<\sum_{j \in I} d_j e_j|\sum_{i \in I} c_i e_i\>$.
+This can be extended to the case where the basis is countable.
 	\end{proof}

### 2020-04-01 JustTheMath/informationGeometry/informationGeometry.tex
Putting together elements of information geometry

\documentclass{article}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
% Remove line spaces between items of enumerate and itemize
\usepackage{enumitem}
\setlist{noitemsep}

\newtheorem{assump}{Assumption}
\renewcommand*{\theassump}{\Roman{assump}}

\newtheorem{axiom}[equation]{Axiom}
\newtheorem{defn}[equation]{Definition}
\newtheorem{prop}[equation]{Proposition}
\newtheorem{coro}[equation]{Corollary}
\newtheorem{thrm}[equation]{Theorem}

%\theoremstyle{definition}

\newenvironment{remark}{\emph{Remark}.}{}
\newenvironment{rationale}{\emph{Rationale}.}{\qed}
\newenvironment{justification}{\emph{Justification}.}{\qed}
\renewenvironment{proof}{\emph{Proof}.}{\qed}

\title{Just the math: information geometry}

\date{\vspace{-5ex}}
\begin{document}

\maketitle

\begin{abstract}
Short summary of information geometry.
\end{abstract}

\section{Parametric model}

\begin{defn}
	We call $\mathcal{S}$ the set of all possible probability density functions.
\end{defn}

\begin{remark}
	We can think of $\mathcal{S}$ as the set of function that are discontinuous at countably many points whose integral is unitary.
\end{remark}

\begin{defn}
	A \textbf{divergence} is a function $D : \mathcal{S} \times \mathcal{S} \to \mathbb{R}$ with the following properties:
	\begin{enumerate}
		\item $D(a, b) \geq 0$ for all $a,b \in \mathcal{S}$
		\item $D(a, b) = 0$ if and only if $a=b$
	\end{enumerate}
\end{defn}

\begin{remark}
	A divergence need not be symmetrical. A divergence need not satisfy the triangle inequality. A divergence is not a distance.
\end{remark}

\begin{defn}
	Given a manifold $M$, a \textbf{continuous parametric model} is a function $P_M : M \to \mathcal{S}$ that returns a probability density function (PDF) of a random variable $X$.
\end{defn}

\begin{defn}
	For a parametric model, we can define the diverge in terms of the point. That is, $D : M \times M \to \mathbb{R}$ where $D[a | b] = D(P_M(a), P_M(b)$ for all $a,b \in M$. Moreover, we can define left and right derivatives as follows:
	\begin{itemize}
		\item $D[\partial_i \cdot | \cdot] : TM \times M \to \mathbb{R}$ such that $D[ (\partial_i) \xi^i(a) | \xi^j(b)] = \partial_i D[\xi^i(a) | \xi^j(b)]$
		\item $D[\cdot | \partial_j \cdot] : M \times TM \to \mathbb{R}$ such that $D[ \xi^i(a) | (\partial_j) \xi^j(b)] = \partial_j D[\xi^i(a) | \xi^j(b)]$
	\end{itemize} 
\end{defn}

%\bibliographystyle{plain}
%\bibliography{bibliography}

\end{document}

### 2020-04-01 Articles/HamiltonianMechanicsIsConservationOfEntropy/refereeResponse3.tex
Preparing for third round of review

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{dutchcal}
\usepackage{braket}
\usepackage{enumitem}

\usepackage{tikz}
\usepackage{forest}
\usetikzlibrary{trees}
\usetikzlibrary{calc}
\usepackage{calculator}
\usepackage{standalone}

\begin{document}

\section*{Reviewer 1 response}

\emph{We again thank reviewer 1 for the continued discussion which we feel helped strengthen the paper.}
\bigskip

\section*{Reviewer 2 response}

\emph{We thank reviewer 2 for the constructive comments and positive recommendation.}
\bigskip

\textbf{I’m unsure of the argument contained in equation (10). If ‘simply a change
	of coordinates’ means $\hat{q} = f(q)$ and $\hat{p} = g(p)$, for some functions $f(\cdot)$, $g(\cdot)$,	then $\{\hat{q}, \hat{p}\} = 1$ implies that $\hat{p} = p/c+A$ and $\hat{q} = cq+B$, for some constants $c$, $A$ and $B$ and the last line in (10) is true only if $A = 0$. On the other
	hand if ‘simply a change of coordinates’ means that $\hat{p} = p/c$ and $\hat{q} = cq$
	then $\{\hat{q}, \hat{p}\} = 1$ is trivially true without (10). Finally if only $\hat{q} = f(q)$ and
	$\{\hat{q}, \hat{p}\} = 1$ are assumed the the fourth line in (10) follows from the third
	only if $\hat{p} = p/c$, in which case it follows that $\hat{q} = cq + B$. }

\emph{By coordinate transformation we mean only $\hat{q} = f(q)$, meaning that the new coordinate $\hat{q}$ is a function of the old coordinate $q$. This is not linear in general, Cartesian to polar being a standard example. With a single coordinate, we could set $\hat{q} = q^3$, then $\hat{p} = \frac{1}{3 q^2} p$. One can verify that $\{\hat{q}, \hat{p}\} = \{ q^3, \frac{1}{3 q^2} p \} = 1$. We modified the text before and after equation (10) to clarify the relationship and highlight the possible non-linearity:}

\bigskip

...

If the change of variables is simply a change of coordinates, then the new coordinate $\hat{q}$ is an arbitrary (possibly non-linear) function of just the old coordinate $q$. We have:
\begin{equation}
\label{coordinate_change}
\begin{aligned}
\hat{q} &= \hat{q}(q) \\
\{\hat{q}, \hat{p}\} &= 1 = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} \\
\dfrac{\partial \hat{p}}{\partial p} &= \frac{\partial \hat{q}}{\partial q} ^{-1} = \frac{\partial q}{\partial \hat{q}} \\
\hat{p} &= \frac{\partial q}{\partial \hat{q}} p
\end{aligned}
\end{equation}
In general, conjugate momentum $\hat{p}$ will depend on the coordinate $q$ through the function $\frac{\partial q}{\partial \hat{q}}$, which is a constant only in the case the transformation is linear. This is exactly the transformation law for the covariant component of a covector. On the other hand, the density and information entropy are invariant:

...

\bigskip

\textbf{I think the last two lines of page 21 and the anecdote in footnote 13 need
	amendment in the light of the paper by Chandrasekar et al. (J. Math.
	Phys. Comm. 48, 032701 (2007) ) who derive forms for the Hamiltonian
	of a damped linear harmonic oscillator.}

\emph{This is a very interesting aspect of mechanics that is a source of great confusion. We were trying to avoid discussing it, since the paper is already long and a suitable treatment would require adequate space. The core issue is that physically different systems can be described by the same set of equations for position/velocity even though the cause for that motion in terms of momentum/energy/force/mass can be different. The best way to address this would be with a few instructive examples that show exactly how a conservative system, through a non-linear coordinate transformation, can be made to look like a non-conservative one. Unfortunately, this would probably add another three or four pages to an already long paper. We added this footnote in the last two lines of page 21, hoping it is enough to give an idea:}

\bigskip

We should again stress that, because every first-order system of equations can be turned into a Hamiltonian one\cite{AllSystemsAreHam}, we can always find a conservative system that in a suitable non-inertial frame ``looks like'' (i.e.~it has the same trajectories of) a non-conservative one. For example, if we see a body coming to a full stop, it could be because drag was acting on it, because it ejected part of its mass, or simply because we accelerated to a comoving frame. The kinematic variables (i.e.~position, velocity and acceleration) are not enough to define whether the system is actually non-conservative, especially if we do not know what reference system we are in. Therefore an equation in terms of $x$, $\dot{x}$ and $\ddot{x}$ is not enough to know we have ``a system under drag''. It is the dynamic variables (i.e.~momentum, energy, force and mass) that provide that characterization: it is the loss of energy through heat that tells us there was drag. Therefore, if one looks closely at Hamiltonian realizations of equations that come from non-conservative systems, such as \cite{chandrasekar2007lagrangian}, one will find non-standard relationships between position, velocity, momentum and energy. At least one of these relationships will have to break down at the equilibrium (e.g.~diverging momentum or energy). There is no way around this because Hamiltonian systems cannot have attractors, while a true dissipative system must have one.

\bibliographystyle{alpha}

\bibliography{bibliography}{}

\end{document}

### 2020-04-01 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
Preparing for third round of review

@@
 \end{equation}
-which we recognize to be the Poisson bracket. If the change of variables is simply a change of coordinates, we have:
+which we recognize to be the Poisson bracket.
+If the change of variables is simply a change of coordinates, then the new coordinate $\hat{q}$ is an arbitrary (possibly non-linear) function of just the old coordinate $q$.
+We have:
 \begin{equation}
@@
 \end{equation}
-We find that conjugate momentum transforms as a covariant component of a covector and that the density and information entropy are invariant:
+In general, conjugate momentum $\hat{p}$ will depend on the coordinate $q$ through the function $\frac{\partial q}{\partial \hat{q}}$, which is a constant only in the case the transformation is linear.
+This is exactly the transformation law for the covariant component of a covector.
+On the other hand, the density and information entropy are invariant:
 \begin{equation}
@@
 Moreover, why should a point-particle be constrained to follow those equations?
-A damped harmonic oscillator does not.
-Why is Hamiltonian mechanics not applicable there?
+A damped harmonic oscillator does not.\footnote{We should again stress that, because every first-order system of equations can be turned into a Hamiltonian one\cite{AllSystemsAreHam}, we can always find a conservative system that in a suitable non-inertial frame ``looks like'' (i.e.~it has the same trajectories of) a non-conservative one.
+For example, if we see a body coming to a full stop, it could be because drag was acting on it, because it ejected part of its mass, or simply because we accelerated to a comoving frame.
+The kinematic variables (i.e.~position, velocity and acceleration) are not enough to define whether the system is actually non-conservative, especially if we do not know what reference system we are in.
+Therefore an equation in terms of $x$, $\dot{x}$ and $\ddot{x}$ is not enough to know we have ``a system under drag''.
+It is the dynamic variables (i.e.~momentum, energy, force and mass) that provide that characterization: it is the loss of energy through heat that tells us there was drag.
+Therefore, if one looks closely at Hamiltonian realizations of equations that come from non-conservative systems, such as \cite{chandrasekar2007lagrangian}, one will find non-standard relationships between position, velocity, momentum and energy.
+At least one of these relationships will have to break down at the equilibrium (e.g.~diverging momentum or energy).
+There is no way around this because Hamiltonian systems cannot have attractors, while a true dissipative system must have one.} Why is Hamiltonian mechanics not applicable there?
 Therefore this characterization does not help us identify the realm of applicability.

### 2020-04-01 Articles/Overview/overview.tex
Adding comments

@@
 
-\extendedtrue
+\extendedfalse
 
@@
 It will contain a summary section at the beginning and a titled paragraph for each of the main ideas or findings.
+
+\ifextended (This extended version contains the same material of the short version, but adds more detail.)
+\else (This short version contains the basic ideas and skips on the details.)
+\fi
 
@@
 If we give a theory, as we said before, by simply stating a mathematical structure, then there is no guarantee that what we give is a fully physical theory.
-Unphysical elements simply means the physical problem was not specified correctly.
+Unphysical elements (e.g.
+mathematical solutions that do not correspond to a physical case) simply means the physical problem was not specified correctly.
 If we truly are able to give a precise physical meaning to every mathematical object, then our formal structures will map one-to-one to our physical understanding.
@@
 \textbf{Physical theories.} This studies the different assumptions that are needed to recover the known physical theories.
-The assumptions typically describe what level of description is available for the system and its parts, how does it change in time and whether trajectories are enough to recover it.
+The assumptions typically describe what level of description is available for the system and its parts, how does it change in time and whether trajectories are enough to recover all information.
 
@@
 This part has not yet been formally developed, though some common ideas are starting to emerge.
-The main idea comes from the realization that, when defining a system, the notions of states, processes, time, evolution laws, interaction with the environment and equilibria are all interrelated and can't be given independently.
+The main idea comes from the realization that, when defining a system, the notions of states, processes, time, law of evolution, interaction with the environment and equilibria are all interrelated and can't be given independently.
 Roughly speaking, to define a system we have to choose a boundary that separates it from the environment.

### 2020-04-01 Articles/2019-Entropy/Source/entropy-article.tex
Reorganized article directory

@@
 
+TODO: Combine with
+
+\section{Introduction}
+
+Cite other axiomatic works.
+
+Single state space to capture:
+* descriptions at all levels
+* descriptions at different times
+* correlations between descriptions at equal time (equations of state)
+* correlations between descriptions at different times (evolution laws)
+* different correlations at different description
+* granularity/precision of such descriptions
+
+We can't have:
+* probability as a native concept
+
+\section{Experimental domains}
+
+First introduce tabular versions of contexts. Introduce equivalence, narrowness, independence and incompatibility on tables.
+
+Introduce verifiable statements and domains. States as possibilities.
+
+Precision over statements.
+
 \bibliography{bibliography}

### 2020-04-01 Notes/2018-IngenuousNonInterpretation/article.tex
Reorganized article directory

\documentclass[letterpaper]{article}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[utf8x]{inputenc}
\usepackage[T1]{fontenc}

%% Sets page size and margins
\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}

% Temporary to remove "submitted to"
\makeatletter
\def\ps@pprintTitle{%
	\let\@oddhead\@empty
	\let\@evenhead\@empty
	\def\@oddfoot{}%
	\let\@evenfoot\@oddfoot}
\makeatother

\usepackage{lineno,hyperref}

\usepackage{amsmath, amsthm, amsfonts}
\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}

\newenvironment{rationale}{\emph{Rationale}.}{\qed}
\newenvironment{justification}{\emph{Justification}.}{\qed}
\renewenvironment{proof}{\emph{Proof}.}{\qed}

\theoremstyle{plain}% default 
\newtheorem{thrm}{Theorem}[section] 
\newtheorem{prop}[thrm]{Proposition} 
\newtheorem{coro}[thrm]{Corollary}

\theoremstyle{definition}
\newtheorem{defn}[thrm]{Definition}
\newtheorem{axiom}[thrm]{Axiom}
\newtheorem*{principle*}{Principle of scientific objectivity}
\theoremstyle{remark}
\newtheorem*{rem}{Remark}

% Remove line spaces between items of enumerate and itemize
\usepackage{enumitem}
\setlist{noitemsep}

% LOGIC symbols
% -------------

% Allows to create negation symbols
\usepackage{MnSymbol}

\DeclareMathOperator{\truth}{truth}
\DeclareMathOperator{\possFn}{poss}
\DeclareMathOperator{\result}{result}
\DeclareMathOperator{\idFn}{id}

\def\TRUE{\textsc{true}}
\def\FALSE{\textsc{false}}

\def\SUCCESS{\textsc{success}}
\def\FAILURE{\textsc{failure}}
\def\UNDEF{\textsc{undefined}}

% Symbols for statements set
\def\stmtSet{\mathcal{S}}
\def\vstmtSet{\mathcal{S}_\textsf{v}}
\def\dstmtSet{\mathcal{S}_\textsf{d}}

% Symbols for tautology and contradiction
\def\tautology{\top}
\def\contradiction{\bot}

% Symbols for "compatibility" and "incompatibility"
\def\comp{\doublefrown}
\def\ncomp{\ndoublefrown}

% Symbols for "narrower" and "wider"
\def\narrower{\preccurlyeq}
\def\nnarrower{\npreccurlyeq}
\def\broader{\succcurlyeq}
\def\nbroader{\nsucccurlyeq}

% Symbol for "independent" and "correlated"
\def\indep{\upmodels}
\def\nindep{\nupmodels}

% Aliases for logical operations
\def\AND{\wedge}
\def\bigAND{\bigwedge}
\def\OR{\vee}
\def\bigOR{\bigvee}
\def\NOT{\neg}

% Formatting for statements
\newcommand{\stmt}[1][s] {\mathsf{#1}}
% Formatting for experimental tests
\newcommand{\expt}[1][e] {\mathsf{#1}}
\newcommand{\exptSet}{\mathcal{E}}
% Formatting for observations
\newcommand{\obs}[1][s] {\mathsf{#1}}
% Formatting for possibilities

% Formatting for experimental domain
\newcommand{\edomain}[1][D] {\mathcal{#1}}

% Formatting for theoretical domain
\newcommand{\tdomain}[1][D] {\bar{\mathcal{#1}}}

\newcommand{\basis}[1][B] {\mathcal{#1}} % Basis

% Formatting for experimental relationships
\newcommand{\erel}[1][r] {#1}

% Formatting for sentence statements
\newcommand{\statement}[1] {\emph{``#1''}}

\begin{document}

\title{The ingenuous non-interpretation of quantum mechanics}
\author{Gabriele Carcassi, Christine A. Aidala \\ University of Michigan}

\date{\today}

\maketitle

\begin{abstract}
	...
\end{abstract}

%\linenumbers

\section{Introduction}

This paper is part of a larger effort, called Assumptions of Physics, which aims at rederiving the known laws of physics from a handful of physical ideas.

Disjointness of physics. Open problems (quantum interpretation, )

The idea is to create a framework We believe that many fundamental problems in physics cannot be solved without first creating a

\section{States and equilibrium}

The first question we ask is: what is the minimum requirement to be able to define the state of a system?

At the very least, the system must exist for a finite amount of time. For example, for a book to be a book the temperature of the surrounding air must be less than 233 Celcius or what we'd have is a pile of ash. While the earth revolves around the sun, the gravitational forces are such that they keep it together. We can study the acceleration of a muon only if it doesn't decay in the meantime.

All physical systems, galaxies, stars, planets, continents, materials, molecules, atoms, fundamental particles, exist in so far that the environment ``let's them be what they are". That is, there is an underlying process, of a faster time scale than our measurements, with which our system is at an equilibrium for at least a finite amount of time.

As states are in equilibrium with such process, if we apply the same process twice we get the same result. If we limit the input for process to be the same space of states it outputs, than if $\mathcal{S}$ is the state space, the process $P : \mathcal{S} \to \mathcal{S}$ is such that $P^2 = P$. That is: it is a projection. For an equilibrium state of that process we have $P(\mathsf{s}) = \mathsf{s}$. That is: it is an eigenstate of that projection operator. So, in order to have states, we also must have projection operators such that the states are eigenstates of that operator.\footnote{Technically, the state space here should be the space of statistical distributions over the possible states. This particular paper aims to give a conceptual understanding of our other more mathematically rigorous work.}

So, without any references to quantum mechanics, observers and the like, we see already the need for projections. These are simply physical processes that have equilibrium. Yes, an experimental physicist may decide to force one system to go through such processes, that's what we do, but the physical process does what it does: projects to an equilibrium state. Why shouldn't we just take the projections in quantum mechanics be these process?

For example, we have an hydrogen atom in equilibrium. We send a photon which is absorbed, making the electron go out of equilibrium. Then it emits a lower energy photon and goes to another state of equilibrium, a higher energy orbital. But that equilibrium is metastable and short lived, the electron emits another photon and goes to the ground state, it's stable equilibrium. Or we have a muon, another metastable equilibrium, which at some point decays into an electron and pair of neutrino/anti-neutrino, a more stable equilibrium.

Since any measurement is, in the end, implemented by a physical process, it's a sequence of emission, decays or scattering of particles. And given that the output of our measurement is going to be the state of something that we can look at over and over, it will also be a projection. That is, we don't need any ad-hoc concept: we just need to realize that a measurement is an equilibrium state of some process.\footnote{Technically, the projection is be defined on the final states which may be different from the states of the system. Mathematically, we can always pull back the projection in the space of states.}

Conceptually, what happens during a projection for a quantum system is the same as when we put hydrogen and oxygen together to make water. The system goes out of equilibrium, so it cannot be described in terms of hydrogen, oxygen and water anymore, then it settles and we can again specify the state of the system based on how much hydrogen, oxygen and water is there. The difference is that in that case we can get a better picture if the state is defined in terms of atoms, which remain what they are during the chemical reaction.

In other words, we can often describe a system out of equilibrium in terms of components which remain withing their own equilibrium. This is something we don't seem to be able to do with quantum systems. Let's look at the relationship between system and subsystems in more detail.

\section{System, subsystems and reducibility}

The second question that we need to explore, then, is: do quantum systems have parts?

Consider a muon: at some point it will decay into an electron and a neutrino-antineturino pair. Consider an electron entering a calorimeter: it will cascade into a shower of electron-positron pairs. In all these cases, everything else stays the same except the original particle is not there anymore and, in its place, there are multiple particles whose total energy is the energy of the original one. The energy/mass of the incoming particle was split into the resulting ones. How would this be possible if the incoming particle wasn't made of parts?

The confusion is between \textit{divisibility}, the idea we have a process such that at the end there are two systems instead of one, and \textit{reducibility}, the idea that at the same time we can describe the system as made of two parts.

Take a planarian flatworm. We can cut into two and from each side a new worm will grow. But a single worm is not made of two worms. That is, the planarian flatworm is divisible into two worms but is not reducible to them. Now take a magnet. We can think of it as made of a north and a south pole. But if we try to break them apart, we get two full magnets. That is, the magnet is reducible into a north and a south pole but is's not divisible into them.

In the same way, we saw that a muon decays into three particle, but it's not made by them. The muon is divisible but not reducible to those particles. Conversely, a proton is thought to be made of quarks and gluons, but we cannot divide it into isolated quarks and gluons. The proton is reducible but not divisible into its constituents.

Therefore there is nothing inconsistent in assuming a system to have parts but not being able to assign them a state. We have cases where whole and parts both have states (e.g. a water molecule and its atoms), cases where only parts have states (e.g. atoms during a chemical reaction) and now cases where only the whole has states (e.g. an electron). Again, we do not need new concepts: just another permutation on the old ones.

If even quantum systems have parts, then, a state will always in some way tell us something about those parts. At the very minimum, if the system is not within a region then neither are its parts. The state of the system may contain the full state of the parts or only a partial description. But it will have to be, in general, some kind of distribution.

The state of a quantum system is, in fact, identified by a wavefunction, which is a type of distribution. What about a classical system?

Again, there may be some confusion here. Some may think that the state of a classical system is defined only by position and momentum, and therefore is not a distribution. Consider the type of systems one describes with classical mechanics: beads, planes, planets. These are not point-like systems: the state can be described just by the position and momentum of the center of mass only in those cases where the physical extent can be ignored. If the orbit of a satellite, for example, would be such that it's distance from the center of the earth is less than its radius, the point particle approximation would fail. That is, in general the state of the system is a distribution $\rho(q^i,p_i)$ in phase space and, when the distribution vanishes outside of a small region where the force is constant, we can just study the center of mass. But this is true for quantum system as well.\footnote{This is Ehrenfest's theorem} Point particles, then, are approximations that can be applied to any sytem.

The parallel between classical states and quantum states is between distributions in phase space and wave functions. But what type of distributions do we have?

This is again a source of confusion as there are three types. The first type describes our propensity to believe that some part has some characteristic. It is subjective in the sense that we may believe different things about the same system. The second type describes the probability to find that some part has some characteristic. It is objective in the sense that the probability is the same for everybody, though it may depend on how the distribution is defined. The third type describes what fraction of the system at that time has some characteristic. It is objective in the sense that it is an actual physical distribution. Which type do we have?

There is actually a tell sign: an element of the distribution of the first two types can be described in isolation. Suppose there is a 50\% probability (or credence) that the system has positive charge, we can study the evolution of the system without worrying about the remaining 50\%. However, if 50\% of the system is positively charged, it will matter whether the remaining 50\% will be neutrally or negatively charged: the trajectories will change.

If we have a classical distribution $\rho(q^i,p_i)$, this will at least self-interact through gravitation. Then it cannot be anything else but an actual physical distribution. If we have a wave-function $\psi(q^i)$, we know that one part may interfere with another and therefore self-interact. Wouldn't it make sense, at this point, to say it must be an actual physical distribution itself? Quite frankly, it seems to us it's the only physically reasonable position.

In this view, which again we believe is not so avoidable, classical and quantum systems are not that different: they are both physical distributions of matter/energy.\footnote{The fact that Hamiltonian mechanics and quantum mechanics have a similar structure is already well known. In our work we show, for example, that classical systems conserve entropy and therefore obey an analogue of the uncertainty principle. There also exists a classical analogue for anti-particles.} We use the same concepts to understand them, the standard ideas of parts and equilibrium we have always had in physics. The only difference is what the distribution can tell us about the parts. A classical system is assumed to be infinitesimally reducible: the state of the whole is equivalent to the states of all infinitesimal parts. A quantum system is irreducible: the state of the whole does not tell us the state of the parts and their internal dynamics.\footnote{In out previous work we showed how these assumptions are enough to recover the mathematical frameork of classical Hamiltonian mechanics on one side and quantum mechanics on the other.}

\section{Quantum systems}

Now that we have seen the basics, we can get to our final question: how are these simple ideas responsible for all quantum effects?

First we note that the inability to characterize the internal structure, the inability to interact with only one part of the system or to prepare it as desired, and the idea that the process under which the quantum system is stable keeps reshuffling its parts are all equivalent: they are restatements of the idea that the system is irreducible. The distribution of the parts remains the same, thus we can define the state for the whole, but the parts keep being reshuffled chaotically, thus we cannot define the state for the parts.\footnote{In classical mechanics, instead, they simply would stay what they were} This is the picture we should have in our head, and with it it becomes straight forward to understand the nature of the different quantum effects.

Uncertainty principle. Here it means that, when we prepare our system, we cannot arbitrarily shrink the distribution in both position and momentum. That is, it's not just a mere uncertainty on what is measure but the inability to prepare a system with a desired precision.\footnote{In both Italian and Spanish it is called indetermination principle, which is more accurate.} This can directly be understood from the fact that we cannot prepare the parts as desired. If we were able to narrow the distribution as we please, we would also be able to control the parts as we please. Another way to see this is that we cannot prepare the parts with better precision than the continuous chaotic reshuffling allows.

Wave particle duality. The wave nature of a quantum corresponds to the distribution of the parts while the particle nature corresponds to the state being defined on the whole. As the state will always be a distribution (e.g. it will be spread in space) so it will always act as a unit (e.g. it cannot be half-absorbed or half-emitted). The idea is that while the motion of each part is chaotic, it is correlated to the motion of all the other parts of the system.

Use of statistics. As the internal dynamics is chaotic, it is natural that statistics plays an important role in quantum mechanics.

Interference patterns. As to parts of the same system start to overlap, so will their chaotic internal dynamics. These will combine as random processes do: depending on their correlation. The more they are correlated, the stronger the combined process. The more they are anti-correlated, the weaker the combined process. Recall how the variance of random variable combines: $\sigma^2_{X+Y} = \sigma^2_{X} + \sigma^2_{Y} + 2 \, \sigma_{X} \sigma_{Y} \rho_{X,Y}$ where $\sigma^2$ is the variance, $\sigma$ the standard deviation, and $\rho_{X,Y}= COV(X,Y)/\sigma_{X}\sigma_{Y}$ is the Pearson correlation coefficient which is a value between $-1$ and $1$. Note how we can use complex numbers to represent the same relationship: $|c_1+c_2|^2=|c_1|^2 + |c_2|^2 + 2 |c_1||c_2|\cos(\arg(c_2) - \arg(c_1))$. The cosine of the phase difference represents the Pearson correlation coefficient.

Exponential decay. If we assume that an isolated quantum particle can decay into multiple ones, we have to assume that such a decay is non-deterministic: we cannot have a one-to-one map between spaces of different dimension. The cause, then, is not just the state of the whole system. As the system is isolated, the only other obvious element that can play a part is the chaotic internal dynamics. Something inside causes to abandon the metastable equilibrium and find a new one, thus the particle decay. Since the internal dynamics is partly to blane, and we have no way of probing the internal dynamics, the probability of decays has to be constant over time. Thus, the time distribution for the decay is exponential.

Non locality. As the system is distributed and as we cannot interact with part of it, interaction with a part of one location is also an interaction on the other location. We can image that while we interact in one location only, the effect on the other location is only mediated through the chaotic internal dynamics, so it cannot establish links of cause and effect, and therefore it cannot be used to transmit information.

Entanglement. As we saw before, to be able to define the state of two systems separately we must have that each is in its own equilibrium. That is, the chaotic internal dynamics of one is independent of the other. Suppose, instead, they are correlated. Then they would act as a single system: they would be entangled. As such, they will display all the characteristics of a single quantum system. When the entanglement is broken, the internal dynamics become independent.

In other words: we can intuitively justify all quantum phenomena. We can do so without introducing any weird or ad-hoc explanation. We are simply talking about

\section{Conclusion}

\section{Acknowledgments}

\bibliography{bibliography}

\begin{thebibliography}{1}

	\bibitem{Turing}
	A. M. Turing: On Computable Numbers, with an Application to the Entscheidungsproblem". Proceedings of the London Mathematical Society 2, 42: 230–265, doi:10.1112/plms/s2-42.1.230, 1936.

	\bibitem{Sipser} M. Sipser: An Introduction to the Theory of Computation, 3rd edition. Cengage Learning, Boston, USA, 2013.

	\bibitem{Shannon} C. E. Shannon: A Mathematical Theory of Communication. The Bell System Technical Journal,
	Vol. 27, pp. 379–423, 623–656, July, October, 1948.

	\bibitem{Pierce} J. R. Pierce: An Introduction To Information Theory: Symbols, Signals and Noise. Dover Books on Mathematics, Second Edition, 1980.

	\bibitem{Brogan} W. L. Brogan: Modern Control Theory, 3rd edition. Pearson, 1990.

	\bibitem{Kalman} R. E. Kalman: A new approach to linear filtering and prediction problems. Journal of Basic Engineering. 82 (1): 35–45, doi:10.1115/1.3662552, 1960.

	\bibitem{Carc1} G. Carcassi, C. A. Aidala, D. J. Baker and L. Bieri: From physical assumptions to classical and quantum Hamiltonian and Lagrangian particle mechanics. Journal of Physics Communications, 2, 4, 045026, 2018.

	\bibitem{Carc2} C. A. Aidala, G. Carcassi, M. J. Greenfield: Topology and experimental distinguishability. arXiv:1708.05492, 2017.

	\bibitem{Carc3} G. Carcassi, C. A. Aidala: Assumptions of Physics (in preparation). http://assumptionsofphysics.org/book/, Ann Arbor, MI, USA, 2018.
\end{thebibliography}

\end{document}

### 2020-04-12 Articles/2019-QuantumComposition/Submissions/NatureComm/tensor.natcom.tex
Version submitted to NatComm

@@
 	which (in contrast to what is done here) are not connected to the
-other postulates. A similar approach was followed in \cite{aerts}
-where tensor products were obtained by specifying some additional
-physical requirements. In quantum field theory one tends to avoid
-problems connected with tensor products of infinite dimensional spaces
-by focusing on algebraic commutation structures, e.g.~\cite{giddins}.
-In particular, the recent MIP*=RE result \cite{mipre} implies that, in
-infinite dimensions, the tensor product is strictly less
-computationally powerful than the commutation structures, emphasizing
-the difference among these two structures, at least for the
-infinite-dimensional case. We will consider the non-relativistic
-setting here.
+	other postulates. Similar approaches appear in \cite{marmo,aerts}
+	where tensor products were obtained by specifying additional physical
+	or mathematical requirements. In quantum field theory one tends to
+	avoid problems connected with tensor products of infinite dimensional
+	spaces by focusing on algebraic commutation structures,
+	e.g.~\cite{giddins,roos}.  In particular, the recent MIP*=RE result
+	\cite{mipre} implies that, in infinite dimensions, the tensor product
+	is strictly less computationally powerful than the commutation
+	structures, emphasizing the difference among these two structures, at
+	least for the infinite-dimensional case. We will consider the
+	non-relativistic setting here.
 
@@
 	whose elements are the two systems and nothing else. We will focus on
-kinematically-independent systems, namely no superselection rules or
+	kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no superselection rules or
 	other restrictions to the state space are present: it is possible to
@@
 	elements in the projective space. We will prove that the map $m$ is
-the tensor product. We focus on pure states here: one can easily
-extend to mixed states using standard tools \cite{nielsenchuang}.
+	the tensor product. We focus on pure states here: the argument can be
+	extended to mixed states using standard tools \cite{nielsenchuang}.
 
@@
 	It has been pointed out before that the quantum postulates are
-redundant: in \cite{masanes} it was shown that the measurement
+	redundant: in \cite{masanes,hartle} it was shown that the measurement
 	postulate (b) can be derived from the others (a), (c), (d). Here
@@
 		Math. (Szeged), {\bf 37}, 263 (1975).
+		\bibitem{marmo} F.M. Ciaglia, A. Ibort, G. Marmo, On the Notion of
+		Composite System, In: Nielsen F., Barbaresco F. (eds) Geometric
+		Science of Information, Lecture Notes in Computer Science, vol
+		11712. Springer (2019).
 		\bibitem{aerts} D. Aerts, I. Daubechies, Physical justification for
@@
 		(2015).% https://doi.org/10.1007/JHEP12(2015)099
+		\bibitem{roos}H. Roos, Independece of Local Algebras in Quantum Field
+		Theory, Comm.  Math. Phys. {\bf 16}, 238 (1970).
 		\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
@@
 			21}, 43 (1991).
-\bibitem{holevo}A. Holevo, Probabilistic and statistical aspects of
-  quantum theory, (North Holland, 1982).
+		\bibitem{hartle}J.B. Hartle, Quantum Mechanics of Individual Systems,
+		Am. J.  Phys. {\bf 36}, 704 (1967).
 	\end{references}

### 2020-04-14 Articles/2019-Entropy/Source/entropy-article.tex
First sketch for process description

@@
 \documentclass[letterpaper]{article}
+
+\input{../../../include/basicstyle}
+\input{../../../include/theorems}
+\input{../../../include/logic}
 
@@
 \section{Introduction}
+
+------------------
+Cite other axiomatic works.
+
+Single state space to capture:
+* descriptions at all levels
+* descriptions at different times
+* correlations between descriptions at equal time (equations of state)
+* correlations between descriptions at different times (evolution laws)
+* different correlations at different description
+* granularity/precision of such descriptions
+
+We can't have:
+* probability as a native concept
+------------------
 
@@
 
-\section{Introduction}
-
-Cite other axiomatic works.
-
-Single state space to capture:
-* descriptions at all levels
-* descriptions at different times
-* correlations between descriptions at equal time (equations of state)
-* correlations between descriptions at different times (evolution laws)
-* different correlations at different description
-* granularity/precision of such descriptions
-
-We can't have:
-* probability as a native concept
-
 \section{Experimental domains}
@@
 
+\section{Process entropy}
+
+The general idea is that we need to describe a system as it evolves in time.
+To that end, we need to define what we mean by states, processes and evolutions in the most general setting such that it can be applied to thermodyamic, dynamical, classical, quantum systems alike.
+
+We call \textbf{process domain} $\tdomain[P]$ the set of all possible descriptions of the system.
+Depending on the system, we will have statements like \statement{the average volume between 1 and 2 seconds is between 3 and 4 liters}, \statement{the half-life of the material is between 4.46 and 4.48 billion years} or \statement{at time 2.5 s the position of the ball is [3.1, 4.2, -1.2] meters}.
+We take these statements to form a logical system, which means $\sigma$-complete Boolean algebra, that is $\tdomain[P]$ is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).
+
+We call \textbf{experimental domain} for the process the subset $\edomain[P] \subseteq \tdomain[P]$ of statements that can be, at least in line of principle, experimentally verified.
+Note that this will not contain, in general, all statements.
+For example, infinite precision statements like \statement{the acceleration is exactly 9.8066543621 $m/s^2$} cannot be verified.\footnote{As discussed further in the appendix, since experimental verification is constrained by finite time and finite resources, $\edomain[P]$ will not be a $\sigma$-complete algebra but, rather, it will be a Heyting algebra: it will only be closed under finite conjunction (AND) and countable disjuction (OR).
+Because our process must be fully defined by physically meaningful statements, we will have that $\tdomain[P]$ will be the closure of $\edomain[P]$ over negation and countable disjunction.
+We also require $\edomain[P]$ to have a countable basis so that the each statement in the domain can, if given enough time, verified.}
+
+We call \textbf{evolution} a statement in $\tdomain[P]$ that, if assumed to be true, will define the truth value for all other statements.
+For example, if our process describes the vertical motion of a point-particle and we posit \statement{the trajectory of the particle is $y=10 \, m - t \cdot 9.80665 \, m/s^2 $} to be true, we would know the truth value of all possible finite and infinite precision measurements at all times.
+A process will therefore define a set of \textbf{possible evolutions}, noted $E \subset \tdomain$, which characterizes what evolutions are physically allowed.
+For example, in a relativistic setting evolutions with velocity greater than the speed of light would not be allowed.
+The possible evolutions, then, will put constraints on the system in terms of logical relationship between statements at different times: if something is true at one time then something else will be true at another time.
+
+It is useful, at this point, to at least mention two types of relationships we can have.
+Given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$ we say one is \textbf{narrower} than the other, noted $\stmt_1 \narrower \stmt_2$, if the second is true whenever the first is.
+This can describe relationships at different scale (e.g.
+\statement{the horizontal position is between 2.5 and 3 meters} $\narrower$ \statement{the horizontal position is between 2 and 3.5 meters}), on different quantities (e.g.
+\statement{the horizontal position is between 2.5 and 3 meters and the vertical position is between 1 and 1.5 meters} $\narrower$ \statement{the horizontal position is between 2.5 and 3 meters}), constraints between variables at the same time (e.g.
+\statement{the temperature of the water in the glass is 3.98 C} $\narrower$ \statement{the density of the water in the glass is 1 $g/cm^3$}) or at different times (e.g.
+\statement{at time 0 s the position is 1 m and the velocity is 1 m/s} $\narrower$ \statement{at time 1 s the position is 2 m}).
+Similarly, we say that two statements are \textbf{compatible}, noted $\stmt_1 \comp \stmt_2$, if they can both be true or incompatible if the cannot.
+This and other relationships can be defined to describe what ultimately will be the equations of state and laws of evolution that govern the process.
+
+For each statement $\stmt \in \tdomain[P]$ we can now define the set $E(\stmt) \subseteq E$ of all the evolutions that are compatible.
+For a point-particle, the evolutions compatible with \statement{at time 0 s the position is 1 m} correspond exactly to all trajectories that pass at the given position at the given time.
+As the process domain $\tdomain[P]$ is a $\sigma$-complete Boolean algebra, the set of all subsets given by $E(\tdomain[P])$ will be $\sigma$-algebra over $E$.\footnote{The experimental domain will instead induce a topology over $E$.
+This can be used as a foundation to extend the framework to include geometrical notions that are outside the scope of this article.} If we are studying a random process, we can think of $E$ as the sample space and $E(\tdomain[P])$ as the event space.
+
+As one can see from the examples, the properties and quantities for the system typically become ways to label the statements.
+Yet, they do not form the basis of the basic structure and, in fact, we will make no use of them.
+All our conclusions, then, will only depend on the idea that we have a process that we study that is describable in terms of experimentally verifiable assertions.
+It does not matter what specific quantities are used, or whether the process is from physics, chemistry, control systems, ecology, economics, sociology, or some other realm we have yet to discover.
+
+\section{States}
+
+As the process extends over time, it is natural to want to organize the statements in our process domain into partial domains, one for each moment in time.
+Given our time parameter $t \in T \subseteq \mathbb{R}$, we can imagine to carve a domain $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
+
+At each moment, we have a set of possible statements about the system corresponding to physically measurable properties.
+We call state the maximal description that can be given at a particular time.
+Given the state, we know all that can be said about the system in that moment.
+We define an evolution of the system as the maximal description of the system at all times.
+If we knew the exact evolution, we know all possible descriptions of the system at all times.
+Therefore picking an evolution is equivalent to picking a state at each time.
+A particular process will constrain how the system can evolve, it will constrain the possible evolutions.
+Therefore we formally define a process as a set of possible evolutions, the ones that are allowed by the process.
+
+As each evolution gives the total description of the system at all times, we assume that all evolutions give the same level of description.
+We also assume we have a way to compare sets of possible evolutions and determine which is bigger and which is smaller.
+Mathematically, we have a measure defined on the set of possible evolutions.
+Each description, if verified experimentally, will narrow the set of possible evolutions to the ones that are compatible with said description.
+Therefore each description, and in particular each state, can be assigned a value that corresponds to the number of evolutions compatible with it.
+We call the logarithm of the number of evolution the entropy associated to the description.
+
+We define a process to be deterministic if knowing the state at one time means we can predict with absolute certainty the state at a future time.
+If this is the case, all evolutions compatible with one state cannot split and must all pass through the same states at all times.
+During a deterministic process, then, the number of evolutions per state can only increase over time.
+The entropy associated to each state is a monotonically increasing function.
+
+Conversely, we define a process to be reversible if knowing the state at one time means we can reconstruct with absolute certainty the state a previous time.\footnote{Note that this notion of reversibility is not identical to the one usually found in thermodynamics, which corresponds to the ``ability to undo the change''.
+We will later find that the notion of thermodynamic reversibility corresponds to the case where system and environment together undergo a deterministic and reversible process.} Reasoning along the same lines, during a reversible process the entropy associated to each state can only decrease over time, as evolutions cannot merge.
+A deterministic and reversible process will necessarily see the number of states, and therefore the entropy, remain the same.
+
+In a deterministic process, a state will be an equilibrium if, roughly speaking, the evolutions have fully merged and they will all give the same future state.
+In other words, after the equilibria the evolution is effectively reversible.
+This means that, at equilibria, the entropy will have reached a maximum.
+Each equilibria will be associated with a bundle of evolutions which will define, at times before the equilibrium, the set of states that will merge.
+For each equilibria one can construct a constraint, an assertion that remains true for each equilibria.
+Therefore each equilibria can be seen as the maximizing the entropy under the constraints imposed by the process.
+
+Finally, a process over a system can be broken into two processes over two systems if and only if they are independent.
+That is, if the evolution over one subsystem tells us nothing about the evolution of the other or, equivalently, the measure on the evolutions on the whole is the product of the measure on the subsystems.
+In those cases, the entropy of the system is equal to the sum of the entropy of the subsystems.
+However, note that if the systems are independent throughout the whole evolution, no coupling can be defined between the subsystem.
+
+A useful scenario is when the systems can be considered independent at the beginning of the evolution, couple, and then return to be independent at equilibrium.
+The entropy can then be considered additive at the beginning and at the end of the evolution, but not in between.
+Equilibria on the whole will necessarily correspond to equilibria on each subsystem.
+Note how this conceptually corresponds to the scenarios studied in thermodynamics.
+
+\section{Math section}
+
+Let $\tdomain$ be the set of all statements about the evolution of the system.
+This is a $\sigma$-complete Boolean algebra.
+Let $E \subset \tdomain$ the set of all evolutions, these are the narrowest statements.
+To each statement $\stmt \in \tdomain$ corresponds a set of evolutions $A(s)$ that are compatible with the statement.
+
+Given two statements $s_1, s_2 \in \mathcal{D}$ we can say whether one is finer than the other $s_1 \finer s_2$.
+Given any statement $\stmt[u] \in \mathcal{D}$, we have a function $\mu_u : \mathcal{D} \to \mathbb{R}^+\cup \{+\infty \}$ such that:
+\begin{enumerate}
+	\item $\mu_u$ is a measure
+	\item $\mu_u(u) = 1$
+	\item $\mu_u(s_1) \leq \mu_u(s_2)$ if $s_1 \finer s_2$
+\end{enumerate}
+The value $\mu_u(\stmt)$ tells us how many trajectories are compatible with $\stmt$ in terms of how many trajectories are compatible with the unit statement $\stmt[u]$.
+
+The conditional probability $P(\cdot | \cdot) : \tdomain \times \tdomain \to [0,1]$ is defined as $P(\stmt_1 | \stmt_2) = \mu_{\stmt_2}(\stmt_1 \AND \stmt_2)$ and represents the fraction of trajectories that are compatible with $\stmt_2$ that are also with $\stmt_1$.
+
+We define $T \subseteq \mathbb{R}$ to be the range of the time parameter.
+We define $\tdomain_{t} \subseteq \tdomain$ as the set of statements defined at the time t.
+Note this does not mean they are statement of only what happens at an instant in time.
+They can be averages around.
+It's how we decompose the problem at the desired time resolution.
+The states at time $t$ are the possibilities of $\tdomain_t$.
+The process is deterministic over the chosen states if for each $x_0 \in \tdomain_{t_0}$ we can find an $x_1 \in \tdomain_{t_1}$ such that $x_0 \narrower x_1$.
+The process is reversible over the chosen states if for each $x_1 \in \tdomain_{t_1}$ we can find an $x_0 \in \tdomain_{t_0}$ such that $x_1 \narrower x_0$.
+
 \bibliography{bibliography}

### 2020-04-16 Articles/2019-Entropy/Source/entropy-article.tex
Another pass at the process domain structure

@@
 To that end, we need to define what we mean by states, processes and evolutions in the most general setting such that it can be applied to thermodyamic, dynamical, classical, quantum systems alike.
+A formal and more rigorous introduction of the framework, both mathematically and conceptually, is given in the appendix.
+Here we limit ourselves to a quick overview of the main point.
+We will we first attempt to use measure theory, which hopefully the reader is already familiar with, show its shortcomings and then present our framework as an amendment.
 
-We call \textbf{process domain} $\tdomain[P]$ the set of all possible descriptions of the system.
-Depending on the system, we will have statements like \statement{the average volume between 1 and 2 seconds is between 3 and 4 liters}, \statement{the half-life of the material is between 4.46 and 4.48 billion years} or \statement{at time 2.5 s the position of the ball is [3.1, 4.2, -1.2] meters}.
-We take these statements to form a logical system, which means $\sigma$-complete Boolean algebra, that is $\tdomain[P]$ is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).
+To define a measure space we start with a set of objects that fully define all variables.\footnote{In probability theory this is called the sample space.} In our case, it will be the set $E$ of \textbf{possible evolutions} which represent full descriptions of the system at all times.
+For example, if our process describes the vertical motion of a point-particle, an evolution would be something like \statement{the trajectory of the particle is $y=10 \, m - t \cdot 9.80665 \, m/s^2 $}.
+Through this set, the process implicitly characterizes what is physically allowed.
+For example, in a relativistic setting evolutions with velocity greater than the speed of light would not be allowed and therefore no be part of $E$.
 
-We call \textbf{experimental domain} for the process the subset $\edomain[P] \subseteq \tdomain[P]$ of statements that can be, at least in line of principle, experimentally verified.
-Note that this will not contain, in general, all statements.
-For example, infinite precision statements like \statement{the acceleration is exactly 9.8066543621 $m/s^2$} cannot be verified.\footnote{As discussed further in the appendix, since experimental verification is constrained by finite time and finite resources, $\edomain[P]$ will not be a $\sigma$-complete algebra but, rather, it will be a Heyting algebra: it will only be closed under finite conjunction (AND) and countable disjuction (OR).
-Because our process must be fully defined by physically meaningful statements, we will have that $\tdomain[P]$ will be the closure of $\edomain[P]$ over negation and countable disjunction.
-We also require $\edomain[P]$ to have a countable basis so that the each statement in the domain can, if given enough time, verified.}
+Next we need a $\sigma$-algebra on $E$ which can be thought as a set of incomplete descriptions about the system.\footnote{In probability theory this is called the set of events.} For a point-particle, the description \statement{at time 0 s the position is 1 m} would correspond to all evolutions such that the particle is at the given position at the given time.
+We call \textbf{process domain} $\tdomain[P]$ the set of all these descriptions as it captures all those descriptions pertinent to the process at hand.
 
-We call \textbf{evolution} a statement in $\tdomain[P]$ that, if assumed to be true, will define the truth value for all other statements.
-For example, if our process describes the vertical motion of a point-particle and we posit \statement{the trajectory of the particle is $y=10 \, m - t \cdot 9.80665 \, m/s^2 $} to be true, we would know the truth value of all possible finite and infinite precision measurements at all times.
-A process will therefore define a set of \textbf{possible evolutions}, noted $E \subset \tdomain$, which characterizes what evolutions are physically allowed.
-For example, in a relativistic setting evolutions with velocity greater than the speed of light would not be allowed.
-The possible evolutions, then, will put constraints on the system in terms of logical relationship between statements at different times: if something is true at one time then something else will be true at another time.
+Finally we a measure $\mu : \tdomain[P] \to \mathbb{R}$ that allows us to give a size to each description.
+This allows us to characterize which description is more or less precise: \statement{at time 0 s the particle is between 0 m and 1 m} is doubly more specific than \statement{at time 0 s the particle is between 0 m and 2 m} precisely because it correspond to half the evolutions of the second.
 
-It is useful, at this point, to at least mention two types of relationships we can have.
-Given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$ we say one is \textbf{narrower} than the other, noted $\stmt_1 \narrower \stmt_2$, if the second is true whenever the first is.
+This measure theoretic view gives a sense of what we would like to do, except it doesn't work and needs to be amended.
+Its biggest strength is its generality.
+Even though we typically use properties and quantities for the system to label the statements, they do not appear in the basic structure.
+Therefore, depending on the system, we will have statements like \statement{the average volume between 1 and 2 seconds is between 3 and 4 liters}, \statement{the half-life of the material is between 4.46 and 4.48 billion years} or \statement{at time 2.5 s the position of the ball is [3.1, 4.2, -1.2] meters} which will correspond to the appropriate set of evolutions.
+The problem is that it fails to handle different level of descriptions in two ways.
+
+First of all, the description given by the evolutions are mathematically different objects from all others: the first are points and the other are sets.
+This is awkward because evolutions of a coarse grain process will be simple descriptions on a fine grain one.
+So we can't properly see one as a mathematical structure contained in the other.
+
+Second, the measure does not have enough resolution: it can only compare objects of the same dimensionality.
+If we imagine all evolutions to be points on an $n$ dimensional manifold, if the measure gives finite values for $k$-dimensional regions it will necessarily give measure zero for all regions with lesser dimensionality and infinite measure for those with greater dimensionality.
+Therefore we will not be able to relate the measure at different level of scale.
+This, again, is a problem since states will have different dimensionality when studying coarse and fine grained descriptions.
+
+To solve the first problem, we note both the evolutions in $E$ and the elements of $\tdomain[P]$ are, in the end, descriptions of the system.
+Therefore we treat as what they are: statements in a well-posed logical system.
+Our process domain $\tdomain[P]$, then, is simply a set of statements that is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND), a $\sigma$-complete Boolean algebra.
+The evolutions $E$ are also descriptions about the system, so we simply have $E \subset \tdomain[P]$.
+All descriptions of subsystems or of the system at lower scale will simply be sub-algebras of $\tdomain[P]$.
+
+All the set operations one would have can be translated in terms of logical operations.
+In the same way that on sets we have an inclusion operator $\subseteq$, given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$ we say one is \textbf{narrower} than the other, noted $\stmt_1 \narrower \stmt_2$, if the second is true whenever the first is.
 This can describe relationships at different scale (e.g.
@@
 \statement{at time 0 s the position is 1 m and the velocity is 1 m/s} $\narrower$ \statement{at time 1 s the position is 2 m}).
-Similarly, we say that two statements are \textbf{compatible}, noted $\stmt_1 \comp \stmt_2$, if they can both be true or incompatible if the cannot.
-This and other relationships can be defined to describe what ultimately will be the equations of state and laws of evolution that govern the process.
+Evolutions, then, can be seen as the narrowest statements in the domain that can still be true.
 
-For each statement $\stmt \in \tdomain[P]$ we can now define the set $E(\stmt) \subseteq E$ of all the evolutions that are compatible.
-For a point-particle, the evolutions compatible with \statement{at time 0 s the position is 1 m} correspond exactly to all trajectories that pass at the given position at the given time.
-As the process domain $\tdomain[P]$ is a $\sigma$-complete Boolean algebra, the set of all subsets given by $E(\tdomain[P])$ will be $\sigma$-algebra over $E$.\footnote{The experimental domain will instead induce a topology over $E$.
-This can be used as a foundation to extend the framework to include geometrical notions that are outside the scope of this article.} If we are studying a random process, we can think of $E$ as the sample space and $E(\tdomain[P])$ as the event space.
+Similarly, we say that two statements are \textbf{compatible}, noted $\stmt_1 \comp \stmt_2$, if they can both be true or incompatible if the cannot, which in terms of set would correspond to the non-empty intersection.
+For each $\stmt \in \tdomain[P]$ we can define $E(\stmt)$ as the set of evolutions that are compatible with it.
+These are exactly the sets we had in the $\sigma$-algebra before, which now corresponds to $E(\tdomain[P])$.
+There are operations, all with straight-forward physical meaning, that can be introduced though we will stop here since these are the only two we strictly require.
 
-As one can see from the examples, the properties and quantities for the system typically become ways to label the statements.
-Yet, they do not form the basis of the basic structure and, in fact, we will make no use of them.
-All our conclusions, then, will only depend on the idea that we have a process that we study that is describable in terms of experimentally verifiable assertions.
+This small change has significant advantages.
+The practical one is that we can now mix and match descriptions at different scale and of different parts of the system within a single framework.
+The conceptual advantage is that it is clear that all we are talking about is descriptions about a system within a model and their logical relationships, which makes it straight-forward to understand our conclusions.
+Lastly, the structure is still very similar to the standard tools of measure theory, therefore, once a scale or subsystem is chose, measure structures can be recovered.
+
+The second problem is solved by replacing the measure with a suitable partial order $\finer$ on $\tdomain[P]$ which, given two statements, it tells us whether the description of one is \textbf{finer}, more refined, than the other.
+Suitable means, for example, that it needs to be compatible with narrowness: if $\stmt_1 \narrower\stmt_2$ we must also have $\stmt_1 \finer \stmt_2$.
+Two statements are \textbf{equigranualar}, noted $\stmt_1 \eqgran \stmt_2$, if they provide the same level of description.
+By analogy with geometrical objects, we can see that these relationships can now span different dimensionality: isolated points are always going to be smaller than lines which are always smaller than surfaces and so on, while we can still have lines and surfaces of the same size.
+For example, we can say that \statement{the trajectory of the particle is $y=t \cdot 1 \, m/s$} $\eqgran$ \statement{the trajectory of the particle is $y=t \cdot 2 \, m/s$} and that \statement{the position at time 0 sec is between 0 and 1 m} $\eqgran$ \statement{the position at time 0 sec is between 1 and 2 m}.
+
+To recover measures, the idea is that, given a statement $\stmt[u] \in \tdomain[P]$ we can construct a measure $\mu_{\stmt[u]}$ such that $\mu_{\stmt[u]}(\stmt[u]) = 1$ and $\stmt_1 \finer \stmt_2$ will mean $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
+That is, we can quantity how more or less precise a statement is compared to a fixed statement taken as a unit.
+
+This does not work in general, though, because statements along different dimensions cannot be compared.
+For example, consider the pressure/volume state space for an ideal gas.
+Comparing \statement{the pressure is 1 kPa and the volume is between 1 and 2 liters} and \statement{the pressure is between 1 and 2 kPa and the volume is 1 liter} would mean saying where one unit liter is bigger or smaller than 1 kPa.
+There are two types of statements in $\tdomain[P]$, though, that can always be compared.
+
+For the first type, note that all evolutions in $E$ must give a complete description, so they must be all equigranular.
+Therefore any statement that corresponds to finitely many evolutions (e.g.
+\statement{the full trajectories is either $y=t \cdot 1 m/s$ or $y=-t \cdot 1 m/s$}) will be comparable to all other statements.
+
+For the second group, note that in our ideal gas example one of the quantities was determined with infinite precision.
+If we gave a range for all quantities, we would compare areas in that space, which we can always do.
+Note that statements with finite precision measurements are the only ones we can experimentally verify.
+So, we will require that all experimentally verifiable statements are comparable.
+We call \textbf{experimental domain} for the process the subset $\edomain[P] \subseteq \tdomain[P]$ of statements that can be, at least in line of principle, experimentally verified.
+
+We leave the full mathematical characterization in the mathematical section.
+We limit ourselves to make a few important comments.
+Verifiable statements will correspond, in terms of sets of $E$, to finite and infinite volumes.
+Requiring those to be comparable means the $E$ is a measurable space.
+In the familiar spaces, such as phase space or physical space, the measure we have will be the familiar one.
+As the process domain corresponds to a $\sigma$-algebra on $E$, the experimental domain corresponds to a topology on $E$.
+Therefore notions of continuity can be recovered as well.
+The process domain is the closure of the experimental domain under standard Boolean operations.
+This means that all the process domain is fully specify by, at least in principle, experimentally verifiable statements.
+The space does not contain any assertion that cannot ultimately be defined in terms of experiments.
+Lastly, requiring that, given enough time, each statement will be verified means the experimental domain must be generated by countably many statements.
+This means the topology is second countable and one is limited to work with ``nice'' spaces (e.g.
+spaces of continuous functions, separable Hilbert spaces, ...) not because they are ``nice'', but because they are the only physically meaningful ones.
+
+To sum up, we have two foundational object.
+The first object is the process domain $\tdomain[P]$ which is a Boolean algebra that contain all the possible description of our process.
+A subset $\edomain[P]$ contains those that are experimentally verifiable, which contain all the information in the full domain.
+The narrowest statements in $\tdomain[P]$ correspond to the evolutions $E$, that provide the full description at all times.
+The second object is the operator $\finer$ which gives us a way to compare two statements and decide which one gives a more refined description.
+Though two statements will not in general be comparable, all verifiable statements are comparable and all evolutions provide an equally refined description.
+
+This framework imposes very few requirements which are of very general applicability.
 It does not matter what specific quantities are used, or whether the process is from physics, chemistry, control systems, ecology, economics, sociology, or some other realm we have yet to discover.
+In the end, if we have a process we need to model, our theory will need at least to provide those objects.
+This means that the results we find will have broad applicability as well.
+
+\section{States}
 

### 2020-04-24 Articles/2019-HamiltonianMechanicsIsConservationOfEntropy/article.tex
Version sent to typesetting

--- 2020-04-01 Articles/HamiltonianMechanicsIsConservationOfEntropy/article.tex
@@
 For example, it is known~\cite{AllHamFreeParticle} that any Hamiltonian system is, at least locally, mathematically equivalent to a free particle.
-That is, we can always find a canonical transformation (i.e.
-a transformation that does not change the equation of motion) such that any Hamiltonian system, in the new coordinates, has the motion of a free particle for a finite amount of time.
+That is, we can always find a canonical transformation (i.e.~a transformation that does not change the equation of motion) such that any Hamiltonian system, in the new coordinates, has the motion of a free particle for a finite amount of time.
 As~\cite{North} notes, only frame-independent relationships should be taken as fundamental.
@@
 
-As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to a recently published work~\cite{AoPPhy1} that aims to identify such starting points.\footnote{We should also mention~\cite{Maudlin} as another attempt to formally derive mathematical structure from physical ideas.} Inspired by those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
+As all these problems stem from a lack of physical principles that can serve as a foundation for classical Hamiltonian and Lagrangian mechanics, we turn to our recently published work~\cite{AoPPhy1} that aims to identify such starting points.\footnote{We should also mention~\cite{Maudlin} as another attempt to formally derive mathematical structure from physical ideas.} Continuing along the lines of those insights, we fully develop another that is only briefly hinted at in that work: the idea that \textbf{a distribution over a manifold conserves information entropy over time if and only if its elements evolve according to Hamiltonian mechanics.} That is, the single requirement of information entropy conservation not only gives us Hamiltonian dynamics but also the structure of phase space (i.e.
 a symplectic structure where variables are organized in conjugate pairs).
@@
 
-We now turn our attention to what~\cite{AoPPhy1} calls an infinitesimally reducible system.
+We now turn our attention to what we call an infinitesimally reducible system~\cite{AoPPhy1}.
 This is a system that can be repeatedly divided into smaller and smaller parts and for which giving a description of the full system is equivalent to giving a full description for all parts.
@@
 In this form $I$ is not a fundamental constant but a parameter that depends upon the initial condition.
-While this is already striking,~\cite{AoPPhy1} shows a simple argument to turn this into something more.
+While this is already striking, one can construct a simple argument to turn this into something more~\cite{AoPPhy1}.
 
@@
 \end{align*}
- This is what~\cite{AoPPhy1} refers to as the classical uncertainty principle.
+ We refer to this as the classical uncertainty principle~\cite{AoPPhy1}.
 
@@
 With this characterization, then, a classical system is one that can be considered to be infinitesimally reducible.
-It is furthermore Hamiltonian if the evolution is deterministic and reversible.\footnote{In the same vein, \cite{AoPPhy1} finds that it is Newtonian if it is purely kinematic (i.e.
-spatial motion identifies the state), it is Lagrangian if it is both Hamiltonian and Newtonian and so on.} The structure of phase space and the equations of motion are a direct consequence of those conditions.
+It is furthermore Hamiltonian if the evolution is deterministic and reversible.\footnote{In the same vein, we find that it is Newtonian if it is purely kinematic (i.e.
+spatial motion identifies the state), it is Lagrangian if it is both Hamiltonian and Newtonian and so on~\cite{AoPPhy1}.} The structure of phase space and the equations of motion are a direct consequence of those conditions.
 It is clear what the realm of applicability of the theory is and its empirical validity is justified within that realm.

### 2020-04-24 Articles/2019-HamiltonianMechanicsIsConservationOfEntropy/arxiv/article.tex
Arxiv version

--- 2020-04-24 Articles/2019-HamiltonianMechanicsIsConservationOfEntropy/article.tex
@@
 \title{Hamiltonian mechanics is conservation of information entropy}
+\author{Gabriele Carcassi*, Christine A. Aidala \\ Physics Department, University of Michigan \\ 450 Church St, Ann Arbor, MI 48109, USA \\
+	Corresponding author: Gabriele Carcassi, carcassi@umich.edu}
 
@@
 \end{abstract}
-
 \maketitle
@@
 
+\section*{Acknowledgements}
+
+We would like to thank Josh Hunt, David J.
+Baker and Gordon Belot for feedback.
+We would especially like to thank Mario Hubert: without his inspiration and guidance this work would have not been possible.
+This article is part of a larger project, Assumptions of Physics, that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.\footnote{See http://assumptionsofphysics.org/book for further details}
+
 \bibliographystyle{alpha}

### 2020-05-24 Proceedings/2019-ICNFP/Carcassi2019ICNFPProceedings.tex
Fixing date
Addressing reviewers comments

@@
 \begin{indented}
-\item[]February 2019
+\item[]February 2020
 \end{indented}
@@
 This is particularly interesting because topological structures are at the foundation of geometrical structures, which play a fundamental role within modern mathematical physics.
-In this talk we will show a set of necessary and sufficient conditions under which those topological structures lead to real quantities and manifolds, which are a typical requirement for geometry.
+In this paper we will show a set of necessary and sufficient conditions under which those topological structures lead to real quantities and manifolds, which are a typical requirement for geometry.
 These conditions will provide a physically meaningful procedure that is the physical counter-part of the use of Dedekind cuts in mathematics.
@@
 For those interested in the ultimate structure of space-time these insights naturally lead to the following line of inquiry.
-Before we want to understand whether space-time is geometrically truly a Riemmanian manifold, we need to understand whether space-time is topologically truly a manifold.
+Before we want to understand whether space-time is geometrically truly a Riemannian manifold, we need to understand whether space-time is topologically truly a manifold.
 A manifold, simply put, is a set of possible cases $X$ that can be experimentally identified by a set of real values.
@@
 The question is when the cases we want to be able to distinguish are parameterized by the real line.
+
+We will concentrate on a single real line (i.e.~total ordering) and not address higher dimensional spaces specifically.
+Since space-time in relativity is four dimensional, it would seem we are ignoring the most interesting cases, but it is not so.
+The issue is that time itself plays a special role: in every local frame, we must be able to use the time coordinate as the affine parameter for the evolution of a particle by writing $x^i=x^i(x^0)$.
+As we mentioned before, functions need to be, at the very least, topologically continuous or they would break the notion of experimental verifiability (i.e.~verifiable statements would not be mapped to other verifiable statements).
+This means that in any theory that includes time evolution, the topology one gives to time will severely constrain the topology of the space within which time evolution takes place.
+Moreover, in any theory of space-time we have the additional constraint that time and spatial coordinates can be mixed.
+Therefore the argument will work in reverse as well: the topology one gives to space has to be the same as the one of time because one may use spatial distance as a time parameter.
+For example, two spaceships drifting at constant velocity may use their spatial separation as a clock.
 
@@
 
-If we think at how quantities are measured, the general idea is that we are able to define references with preset amounts, and then compare our object to several references to find the ones that bound the value from above and below.
+If we think about how quantities are measured, the general idea is that we are able to define references with preset amounts, and then compare our object to several references to find the ones that bound the value from above and below.
 For example, a ruler is a series of marks and measuring the position with it means finding the closest one.
@@
 These are the sort of basic ideas and language we use within our framework, and from these we work out a set of necessary and sufficient conditions that a set of references needs to satisfy so that they define a set of possible distinguishable cases, or possibilities for short, that is ordered like the real line.
+
+The idea is that the quantities themselves are not a priori objects, but a construction built upon a set of references that form our system of measurement.
+A reference frame in space-time, then, is also a construction consisting of fixed elements (e.g.
+the stars, the borders of the experiment table, the elements of a timing system in a particle accelerator, ...) and signals exchanged among them.
+The topology of space-time is then an idealized characterization of the set of all possible such constructions, that is, all possible reference systems.
+Our goal is to understand the extent and the limitation of such idealization.
 
@@
 The first major finding is that \emph{most of the conditions are not required by the real numbers specifically, but they are required by ordered quantities in general}.
-That is, the set of reference has to satisfy a few conditions just to be sure that the cases they distinguish experimentally are ordered (i.e.~can be thought to be all being one after the other).
+That is, the set of references has to satisfy a few conditions just to be sure that the cases they distinguish experimentally are ordered (i.e.~can be thought to be all being one after the other).
 The conditions are the following.
@@
 In this light, the ultimate structure of space-time may not be geometrical but it will still be topological, as we will still need to describe what is experimentally accessible.
+At large scale, geometrical structure will need to emerge to recover the established theories.
+The point is that these large scale geometric structures cannot emerge from other geometrical structures, as these would necessarily suffer from the same fundamental problems.
 
@@
 
+To give a better understanding of how these definitions work in a multidimensional setting, let us see how they apply to a Cartesian frame with coordinates $(x, y, z)$.
+A reference for the $x$ coordinate would be something like $r_1=$(``$x < 0$'',``$0 \leq x \leq 2$'', ``$ x > 2$'').
+Note that it partitions the whole space in three disjoint regions.
+The on case acts as a divider and extends throughout $y$ and $z$.
+The reference $r_2=$(``$y < 0$'',``$0 \leq y \leq 1$'', ``$y > 1$'') is not aligned with $r_1$ as something can be before $r_1$ and still be before, on or after $r_2$.
+On the other hand, $r_3=$(``$x < -1$'',``$-1 \leq x \leq 1$'', ``$x > 1$'') is aligned with $r_1$ even though it is not either before or after $r_1$ because the on regions overlap.
+Note that these references would not have such a simple expression in, say, spherical coordinates.
+Yet the regions themselves are coordinate independent and, therefore, the relationships between them are too.
+
+What happens is that each reference frame would have its own set of references that can be nicely expressed using its coordinates.
+The before/after relationships expressed using one set of references may not be simply expressible using another, precisely because they are not aligned with each other.
+Relativistically, the references used for time by one frame will not be aligned with the ones of a boosted frame: the coordinate surfaces do not partition (i.e.
+do not foliate) space-time in the same way.
+Reference alignment essentially captures these various requirements with one simple formal definition.
+
 \subsection{Refinable references}
@@
 
-What does this requirement entail in practice? It means that we can place a reference wherever we want and can shrink its extent such that it occupies only one possibility.
+What does this requirement entail in practice?
+It means that we can place a reference wherever we want and can shrink its extent such that it occupies only one possibility.
+If we take the traditional manifold structure of space-time literally, this would mean having at our disposal ever shrinking references that can be placed anywhere in space and time.
+For example, it would require being able to create a timing system with as many well separated pulses as desired distributed to as many places as desired.
+All of this, without changing the nature of the process we are studying.
 
@@
 The inclusion of all the limit points will come from the infinite union allowed by the topology while the dense subset is associated with the countable base, which is a requirement for our topology to be physically meaningful.
+
+%Note that references have to be physically implemented somehow, either by placing massive objects or sending signals at the reference points, which will affect the object under study, at the very least by gravitational effects.
+But for many applications such reference-induced effects can be neglected.
+and the process can be repeated sufficiently well that, in practice, we can assume
+
+%Riemannian
+%set of reference (missing s)
+%think at
 

### 2020-06-16 Articles/2019-Entropy/Source/entropy-article.tex
More work on the thermo foundations

--- 2020-04-14 Articles/2019-Entropy/Source/entropy-article.tex
@@
 
-\section{Different level of precision}
+\section{Main points}
+\begin{enumerate}
+	\item Entropy assigns a size to points. Shannon entropy works on a distribution where each point has the same entropy.
+	\item Measure theory differentiate between points and sets. This does not work when combining different level of granularity.
+\end{enumerate}
 
-We regard states as points in a space. Conceptually, this presents problems. Mathematically, what we care about is the sigma algebra.
+\section{Main section}
+\begin{enumerate}
+	\item Model of a process.
+Set of all possible descriptions at all times.
+Some descriptions are narrower than others.
+Evolutions are the narrowest.
+Some descriptions are compatible.
+Each description has a unique set of compatible evolutions.
+Descriptions in narrower if the set of compatible evolution is a subset.
+	\item We can group those descriptions relative at a particular time. It is actually interval of time. States are the narrowest description at that time. How are states related through the evolution?
+	\item Define deterministic evolution: can predict future state.
+Past state narrower.
+Number of evolutions increases.
+Define reversible.
+Future state narrower.
+Number of evolution decreases.
+Det/rev remains the same.
+How do we quantify that?
+	\item Define way to compare granularity. Not a single measure, not everything comparable or finitely comparable. Define probability in terms of measure.
+	\item States of the same system must be finitely comparable.
+Can put a measure on states themselves, which is a count of evolutions.
+Examples to show that: the measure and state space, in general, change over time.
+State spaces are stable and can be assigned a measure only at equilibrium.
+	\item Define independent systems as those where the state of one does not constrain the state of the other. The measure must be the product.
 
-Introduce theoretical domains as point-less spaces, descriptions, that follow a countable boolean algebra. A size is given by a measure, additive map from the theoretical domain to the reals.
+\end{enumerate}
 
-To define a unit system, we take one statement to be the unit, the size of a set can be defined as the ratio of the measure of the set and the measure of the set of the unit.
+\begin{enumerate}
+	\item A thermodynamic system is one that undergoes deterministic evolution with equilibria. There are variables that  
+	\item Extensive quantities are somewhat postulated at this point. So is the existence of energy.
+	\item Second law come from deterministic system.
+Define entropy as the log of the measure, so that it is additive under independent systems.
+We have: entropy always increases under deterministic evolution; entropy is maximal at equilibrium; entropy is additive for independent system.
+Entropy is a function of the equilibrium, which is identified by state variables.
+	\item Third law comes from states of the same system being finitely comparable. Lowest entropy is "one evolution" in terms of system granularity.
 
-Two statements are independent if all minterms are possible, are not contradictions.
+\end{enumerate}
 
-TODO: Combine with
+\section{Evolutions, states and levels of description}
 
-\section{Experimental domains}
+The overall goal is to describe a generic system as it evolves in time by defining a minimal set of concepts that such a setup must define.
+We want to find notions that can be applied, for example, to thermodyamic and dynamical systems alike.
+Then we want to derive general results based on such definitions.
+Here we provide an overview of the main features, leaving the formal definitions in the (yet to be written) appendix.
 
-First introduce tabular versions of contexts. Introduce equivalence, narrowness, independence and incompatibility on tables.
+\subsection{Descriptions and logical relationships}
 
-Introduce verifiable statements and domains. States as possibilities.
+Our starting point is the \textbf{process domain} $\tdomain[P]$, which is the set that contains all possible descriptions of the system at all levels of detail at all possible times.
+Depending on the process, it will contain statements like \statement{the average volume between 1 and 2 seconds is between 3 and 4 liters} or \statement{the trajectory of the particle is $y=10 \, m - t \cdot 9.80665 \, m/s^2 $}.
+It does not matter at this point what the statements themselves are, except they must be, at least in line of principle, defined from experimentally well-defined starting points.
+Mathematically, $\tdomain[P]$ is a $\sigma$-complete Boolean algebra\footnote{A set of statements that is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).} and it is the closure of the set of experimentally verifiable statements $\edomain[P] \subseteq \tdomain[P]$.\footnote{These will form a Heyting algebra which, to be experimentally reachable, must allow a countable basis.}
 
-Precision over statements.
-
-\section{Process entropy}
-
-The general idea is that we need to describe a system as it evolves in time.
-To that end, we need to define what we mean by states, processes and evolutions in the most general setting such that it can be applied to thermodyamic, dynamical, classical, quantum systems alike.
-
-We call \textbf{process domain} $\tdomain[P]$ the set of all possible descriptions of the system.
-Depending on the system, we will have statements like \statement{the average volume between 1 and 2 seconds is between 3 and 4 liters}, \statement{the half-life of the material is between 4.46 and 4.48 billion years} or \statement{at time 2.5 s the position of the ball is [3.1, 4.2, -1.2] meters}.
-We take these statements to form a logical system, which means $\sigma$-complete Boolean algebra, that is $\tdomain[P]$ is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).
-
-We call \textbf{experimental domain} for the process the subset $\edomain[P] \subseteq \tdomain[P]$ of statements that can be, at least in line of principle, experimentally verified.
-Note that this will not contain, in general, all statements.
-For example, infinite precision statements like \statement{the acceleration is exactly 9.8066543621 $m/s^2$} cannot be verified.\footnote{As discussed further in the appendix, since experimental verification is constrained by finite time and finite resources, $\edomain[P]$ will not be a $\sigma$-complete algebra but, rather, it will be a Heyting algebra: it will only be closed under finite conjunction (AND) and countable disjuction (OR).
-Because our process must be fully defined by physically meaningful statements, we will have that $\tdomain[P]$ will be the closure of $\edomain[P]$ over negation and countable disjunction.
-We also require $\edomain[P]$ to have a countable basis so that the each statement in the domain can, if given enough time, verified.}
-
-We call \textbf{evolution} a statement in $\tdomain[P]$ that, if assumed to be true, will define the truth value for all other statements.
-For example, if our process describes the vertical motion of a point-particle and we posit \statement{the trajectory of the particle is $y=10 \, m - t \cdot 9.80665 \, m/s^2 $} to be true, we would know the truth value of all possible finite and infinite precision measurements at all times.
-A process will therefore define a set of \textbf{possible evolutions}, noted $E \subset \tdomain$, which characterizes what evolutions are physically allowed.
-For example, in a relativistic setting evolutions with velocity greater than the speed of light would not be allowed.
-The possible evolutions, then, will put constraints on the system in terms of logical relationship between statements at different times: if something is true at one time then something else will be true at another time.
-
-It is useful, at this point, to at least mention two types of relationships we can have.
-Given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$ we say one is \textbf{narrower} than the other, noted $\stmt_1 \narrower \stmt_2$, if the second is true whenever the first is.
-This can describe relationships at different scale (e.g.
-\statement{the horizontal position is between 2.5 and 3 meters} $\narrower$ \statement{the horizontal position is between 2 and 3.5 meters}), on different quantities (e.g.
+As $\tdomain[P]$ contains different levels of description, one statement may give a more specific description than another.
+For example, we say that \statement{the horizontal position is between 2.5 and 3 meters} is \textbf{narrower} than \statement{the horizontal position is between 2 and 3.5 meters}.
+Note that whenever the first statement is true, the second one must be true as well.
+Given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$, we formally capture this relationship by noting $\stmt_1 \narrower \stmt_2$.
+Narrowness can describe relationships at different scales, on different quantities (e.g.
 \statement{the horizontal position is between 2.5 and 3 meters and the vertical position is between 1 and 1.5 meters} $\narrower$ \statement{the horizontal position is between 2.5 and 3 meters}), constraints between variables at the same time (e.g.
@@
 \statement{at time 0 s the position is 1 m and the velocity is 1 m/s} $\narrower$ \statement{at time 1 s the position is 2 m}).
-Similarly, we say that two statements are \textbf{compatible}, noted $\stmt_1 \comp \stmt_2$, if they can both be true or incompatible if the cannot.
-This and other relationships can be defined to describe what ultimately will be the equations of state and laws of evolution that govern the process.
+It is a general tool to capture relationships within the theory which intuitively can also be thought as implication.\footnote{Technically, implication in classical logic is a truth function as classical logic does not incorporate the idea of different ``possible cases''.
+What we have is more similar in spirit to semantic consequence in modal logic, though modal logic brings in a lot of undesirable elements we do not want.}
 
-For each statement $\stmt \in \tdomain[P]$ we can now define the set $E(\stmt) \subseteq E$ of all the evolutions that are compatible.
-For a point-particle, the evolutions compatible with \statement{at time 0 s the position is 1 m} correspond exactly to all trajectories that pass at the given position at the given time.
-As the process domain $\tdomain[P]$ is a $\sigma$-complete Boolean algebra, the set of all subsets given by $E(\tdomain[P])$ will be $\sigma$-algebra over $E$.\footnote{The experimental domain will instead induce a topology over $E$.
-This can be used as a foundation to extend the framework to include geometrical notions that are outside the scope of this article.} If we are studying a random process, we can think of $E$ as the sample space and $E(\tdomain[P])$ as the event space.
+Within our process domain $\tdomain[P]$ we can define the set $E$ of the narrowest possible statements.
+Its elements give a complete description of our system at all times and therefore we call them \textbf{evolutions} while we call $E \subset \tdomain[P]$ the set of all possible evolutions.
 
-As one can see from the examples, the properties and quantities for the system typically become ways to label the statements.
-Yet, they do not form the basis of the basic structure and, in fact, we will make no use of them.
-All our conclusions, then, will only depend on the idea that we have a process that we study that is describable in terms of experimentally verifiable assertions.
+Another logical relationship we want to capture is the idea of whether two statements can be true at the same time.
+For example, we want to say that \statement{the horizontal position is between 2.5 and 3.5 meters} is \textbf{compatible} with \statement{the horizontal position is between 2 and 3 meters}.
+In general, given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$ we note $\stmt_1 \comp \stmt_2$ if it is possible for them to be both true.
+As with narrowness, compatibility can be in terms of different ranges of the same quantities, of different quantities and of constraints at equal or different times.
+
+Every statement $\stmt \in \tdomain[P]$ can now be characterized by the set $A(\stmt) \subseteq E$ of all evolutions compatible with $\stmt$, that is all the evolutions for which $\stmt$ will be true.
+Mathematically,  $A(\edomain[P])$ is a topology over $E$ while $A(\tdomain[P])$ forms a $\sigma$-algebra over $E$, the Borel algebra of $A(\edomain[P])$.
+Narrowness and compatibility become set realtionships in the $\sigma$-algebra: $\stmt_1 \narrower \stmt_2$ if and only if $A(\stmt_1) \subseteq A(\stmt_2)$ (i.e.~all trajectories for which $\stmt_1$ is true are such that $\stmt_2$ is also true) and $\stmt_1 \comp \stmt_2$ if and only if $A(\stmt_1) \cap A(\stmt_2) \neq \emptyset$ (i.e.~two descriptions are compatible if there is an evolution that is compatbile with both).
+
+\begin{table}[h!]
+	\centering
+\begin{tabular}[h]{|c|c|p{6cm}|}
+	\hline 
+	Symbol & Name & Meaning \\ 
+	\hline 
+	$\tdomain[P]$ & Process domain & The set of all possible descriptions at all times and at all levels of granularity \\ 
+	\hline 
+	$E \subset \tdomain[P]$ & Possible evolutions & The set of descriptions that give a full account of the process; they correspond to the narrowest statements within $\tdomain[P]$ \\ 
+	\hline 
+	$A(\stmt)$ & Compatible evolutions & The set of evolutions that are compatible with the statement $\stmt$ \\ 
+	\hline 
+	$\narrower$ & narrower & A statement is narrower than another $\stmt_1 \narrower \stmt_2$ if the second one is true whenever the first one is \\ 
+	\hline 
+	$\comp$ & compatible & A statement is compatible with another $\stmt_1 \comp \stmt_2$ if they can be true at the same time \\ 
+	\hline 
+\end{tabular} 
+	\caption{Process descriptions and their logical relationships}
+	\label{table:logic}
+\end{table}
+
+\subsection{States and time}
+
+As the process extends over time, we want to organize the statements in our process domain into partial domains, one for each moment in time.
+Given a \textbf{time parameter} $t \in T \subseteq \mathbb{R}$, we can imagine to carve a subdomain $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
+Note that $\tdomain_t$ is a $\sigma$-complete Boolean sub-algebra of $\tdomain[P]$ since any logical operation on statements at fixed $t$ is also another statement at fixed $t$.
+The set $X_t \subset \tdomain_t$ of the narrowest statements will give the most precise description at that time: these are the possible states of our system at that time.
+
+We should note that, technically, the statements are not defined at an infinitesimal moment $t$, rather within a finite interval $[t, t + \Delta t]$.
+This represents the time-scale at which the process will be described, meaning the description within $\tdomain_t$ should not be sensitive to what happens at faster scales.
+The quantities used to represent states at time $t$, then, should really be thought as averages within $\Delta t$.
+In the same way, $\tdomain_t$ will in general represent the system at the chosen level of description.
+Statements about the environment or at a finer level of description (i.e.
+the positions of all molecules for a gas) will not be included.
+
+We should also note that the set of states $X_t$, in general, may not be the same at all times.
+In fact $X_t$ will only include ``possible states'', those that are visited at least by one possible evolution (i.e.
+$A(x) \neq \emptyset$ for all $x \in X_t$), and it may change in time.
+For example, the set of possible states for a dissipative system will shrink as they converge to equilibiurm.
+In quantum mechanics, the possible states after a measurement are only the eigenestates and a different choice of measurement will lead to a different process with different possible evolutions.
+The idea of a state space of the system, independent of the process, is something we need to construct.
+Finding the requirements needed for that construction, the hidden assumptions, is precisely the issue we need to address.
+
+We say a process is \textbf{deterministic} over the time domains $\{X_t\}_{t \in T}$ if given the state at one time we can always predict the state at all future times.
+Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_1 \in X_{t_1}$ we can find an $x_2 \in X_{t_2}$ such that $x_1 \narrower x_2$.
+
+Intuitively, all the evolutions that pass through $x_1$ will also pass through $x_2$.
+If $x(t)$ is the state trajectory in time, we have $A(x(t)) \subseteq A(x(t + \Delta t))$: the set of compatible trajectories must stay the same or become larger.
+
+Conversely, we say a process is \textbf{reversible} over the time domains $\{X_t\}_{t \in T}$ if given the state at one time we can always reconstruct the state at all past times.
+Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_2 \in X_{t_2}$ we can find a $x_1 \in X_{t_1}$ such that $x_2 \narrower x_1$.
+
+In this case, all the evolutions that pass through $x_2$ must have passed through $x_1$.
+If $x(t)$ is the state trajectory in time, we have $A(x(t)) \subseteq A(x(t - \Delta t))$: the set of compatible trajectories must stay the same or become smaller.
+
+A deterministic and reversible process, then, will mean $x_1$ and $x_2$ are equivalent and that  $A(x(t)) = A(x(t + \Delta t))$: the set of evolutions remains the same.
+The fact that past and future states are equivalent $x(t) \equiv x(t + \Delta t)$ does not mean they are the same statement: it means there is an if-and-only-if relationship between them.
+For example, \statement{at time $t$ the position is $q$ and the velocity is $v$} if and only if \statement{at time $t + \Delta t$ the position is $q + v \Delta t$ and the velocity is $v$}.
+The differential equation $\dot{v} = 0$ is a short hand for all such relationships.
+
+If we limit ourselves to deterministic procseses, with these few definitions, we have already found something that ``grows bigger'' under irreversible evolution while it remains the ``the same size'' under reversible evolution.
+
+TODO: picture with deterministic evolution merging "streams" of evolutions, and detrev evolutions not merging
+
+\begin{table}[h!]
+	\centering
+	\begin{tabular}[h]{|c|c|p{6cm}|}
+		\hline 
+		Symbol & Name & Meaning \\ 
+		\hline 
+		$t \in T$ & Time parameter & Time is treated as a real valued parameter within the range $T \subseteq \mathbb{R}$ \\ 
+		\hline 
+		$\tdomain_t \subset \tdomain[P]$ & Time domain & The set of descriptions at a given time \\ 
+		\hline 
+		$X_t \subset \tdomain_t$ & Possible states & The set of descriptions that give a full account at time $t$; they correspond to the narrowest statements within $\tdomain_t$ \\ 
+		\hline 
+		$x(t) \narrower x(t + \Delta t)$ & Determinism & Given the state at one time we can predict the state at future times \\ 
+		\hline 
+		$x(t) \narrower x(t - \Delta t)$ & Reversibility & Given the state at one time we can reconstruct the state at past times \\ 
+		\hline 
+	\end{tabular} 
+	\caption{Time domain, states and determinism}
+	\label{table:states}
+\end{table}
+
+\subsection{Granularity and quantifying evolutions}
+
+Next step is to quantify the size of the sets of evolutions, specifically those that correspond to statements.
+Since statement at a finer level of description will put a greater constrain to the set of evolutions, quantifying the size of $A(\stmt)$ means quantifying the granularity of the description of $\stmt$.
+The statement \statement{at time 0 s the particle is between 0 m and 1 m} is doubly more specific than \statement{at time 0 s the particle is between 0 m and 2 m} precisely because it correspond to half the evolutions of the second statement.
+
+The naive approach would be to assign a measure to sets of $E$ and since $A(\tdomain[P])$ is a $\sigma$-algebra, it would seem we are done.
+Unfortunately it is not that simple: a single measure can only compare objects of the same dimensionality.
+If we imagine all evolutions to be points on an $n$ dimensional manifold, if the measure gives finite values for $k$-dimensional regions it will necessarily give measure zero for all regions with lesser dimensionality and infinite measure for those with greater dimensionality.
+Coarse grained and subsystem descriptions will have different dimensionality therefore we need a way to quantify objects at all levels.\footnote{Typically one uses geometrical structure for his purpose.
+In phase space, the symplectic form allows us to define areas of two-dimensional surfaces, and use those to define volumes or even-dimensional areas.
+Note that it does not quantify the size of the points, which are implicitly assumed to be of equal size.
+This is a special case and cannot be the starting point.}
+
+Therefore we will assume that we have a way to compare two statements and decide which one gives a refined description of the process.
+Formally, we have partial order $\finer$ on $\tdomain[P]$ which, given two statements, it tells us whether the description of one is \textbf{finer}, more refined, than the other.
+Two statements are \textbf{equigranualar}, noted $\stmt_1 \eqgran \stmt_2$, if they provide the same level of description.
+For example, we can say that \statement{the trajectory of the particle is $y=t \cdot 1 \, m/s$} $\eqgran$ \statement{the trajectory of the particle is $y=t \cdot 2 \, m/s$} and that \statement{the position at time 0 sec is between 0 and 1 m} $\eqgran$ \statement{the position at time 0 sec is between 1 and 2 m}, even though the second pair is infinitely less discerning than the first.
+Note that if one statement is narrower than the other (i.e.
+$\stmt_1 \narrower \stmt_2$) then it is also finer than the other (i.e.
+$\stmt_1 \finer \stmt_2$) while the reverse is not necessarily true.
+
+We say two statements are \textbf{comparable} if one is finer than the other.
+The order is partial because not all statements are comparable to each other.
+For example, consider the pressure/volume state space for an ideal gas.
+Comparing \statement{the pressure is 1 kPa and the volume is between 1 and 2 liters} and \statement{the pressure is between 1 and 2 kPa and the volume is 1 liter} would mean saying where one unit liter is bigger or smaller than 1 kPa.\footnote{In phase space, for example, we cannot in general compare $q^i$ and $p_i$, yet we can always compare areas of phase space.}
+
+To recover measures, we pick a statement $\stmt[u] \in \tdomain[P]$ and construct a \textbf{measure} $\mu_{\stmt[u]}$ such that $\mu_{\stmt[u]}(\stmt[u]) = 1$ and $\stmt_1 \finer \stmt_2$ will mean $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
+That is, we can quantity how more or less precise a statement is compared to a fixed statement taken as a unit.
+If two statements are such that the measure of one with respect to the other is finite and non-zero, we say they are \textbf{finitely comparable}.
+Furthermore, since all evolutions give a complete description of the whole process, we will assume they are all equigranular.
+Therefore if we pick $\stmt[u] \in E$, then $\mu_{\stmt[u]}(\stmt)=\#(A(\stmt))$ would be the counting measure of evolutions compatible with $\stmt$.\footnote{We still needs to understand the necessary and sufficient conditions under which such measures would exist and be unique.
+For the present work, we assume it can be done.
+We know that it is a necessary condition that all evolution must be equigranular.
+It remains an open question is whether it is sufficient.}
+
+We can also define the \textbf{dynamical entropy} with respect to $\stmt[u]$ the quantity $S_{\stmt[u]}(\stmt) = \log \mu_{\stmt[u]}(\stmt)$.
+This can be thought as the number of bits, yes/no questions, that separate the level of description of $\stmt[u]$ from the one provided by $\stmt$.
+Connecting the definitions from above, under a deterministic process we have $S_{\stmt[u]}(x(t)) \leq S_{\stmt[u]}(x(t + \Delta t))$ since $x(t)\narrower x(t+\Delta t)$ means $x(t) \finer x(t+\Delta t)$ and $\mu_{\stmt[u]}(x(t)) \leq \mu_{\stmt[u]}(x(t+\Delta t))$.
+If the process is determistic and reversible, we have $S_{\stmt[u]}(x(t)) = S_{\stmt[u]}(x(t + \Delta t))$.
+In other words, \textbf{dynamical entropy cannot decrease for a deterministic process and is stationary if the process is reversible as well}.
+
+Quantifying evolutions allows us to define probability as well.
+If we consider all the evolutions compatible with \statement{the coin is fair}, we must have that half of them are compatible with \statement{the coin is head} while the other are compatible with \statement{the coin is tails}.
+Therefore we have $P(\stmt_2 | \stmt_1) = \mu_{\stmt_1}(\stmt_1 \AND \stmt_2)$ where $\stmt_1 \AND \stmt_2$ is the logical conjunction (AND) between the two statements.
+In other words, the probability quantifies the ratio of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$.
+
+%Note that this approach resolves a problem that causes friction between quantum mechanics and standard probability theory.
+Given two incompatible states $\psi$ and $\phi$, that is $0 < | \left<\psi | \phi \right> |^2 < 1$, we would like to say that that the probability of one given the other is $P(\psi | \phi) = | \left<\psi | \phi \right> |^2$.
+
+\begin{table}[h!]
+	\centering
+	\begin{tabular}[h]{|c|c|p{6cm}|}
+		\hline 
+		Symbol & Name & Meaning \\ 
+		\hline 
+		$\finer$ & finer & A statement is finer than another $\stmt_1 \finer \stmt_2$ if it provides a more refined description than the second \\ 
+		\hline 
+		$\eqgran$ & equigranular & A statement is equigranular to another $\stmt_1 \eqgran \stmt_2$ if they provide the same level of detail \\ 
+		\hline 
+		$\mu_{\stmt[u]}(\stmt)$ & Measure & Quantifies the precision of a statement $\stmt$ in terms of a reference unit statement $\stmt[u]$; it quantifies how many evolutions are compatible with $\stmt$ compared to the evolutions compatible with $\stmt[u]$  \\
+		\hline 
+		$S_{\stmt[u]}(\stmt)$ & Dynamical entropy & The number of bits that separate the level of description of $\stmt$ from the one provided by the unit statement $\stmt[u]$; it quantifies the number of questions needed to go from the level of description of $\stmt$ to the level of description of $\stmt[u]$  \\
+		\hline 
+		$P(\stmt_2 | \stmt_1)$ & Probability & Quantifies quantifies the ratio of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$  \\ 
+		\hline 
+		 & Comparable & Two statements are comparable if one is finer than the other \\ 
+		\hline 
+		 & Finitely comparable & Two statements are finitely if the measure of one in terms of the other is finite and non-zero \\ 
+		\hline 
+	\end{tabular} 
+	\caption{Granularity, measures and probability}
+	\label{table:states}
+\end{table}
+
+\subsection{Systems, equilibria and state spaces}
+
+So far we have only characterized how descriptions at different times can be related to each other.
+We now want to group descriptions that belong together, that describe a single system as it is subjected to different processes.
+In the simplest form, we prepare the system in an initial state, let it evolve and look at the final state.
+We want to characterize the state space of all possible initial and final states under all process and characterize the process.
+Our framework already tells us that if the process is deterministic, the process can be characterized as a map $x(t+\Delta t) = f(x(t))$.
+If the process is non-deterministic, the evolutions for each state may split, and the probability $P(x(t + \Delta t) | x(t))$ can be used to characterize the fraction of evolutions that start in $x(t)$ and end in $x(t+\Delta t)$.
+Yet, it also tells us there are a few things that we have to worry about.
+
+First of all, the process can't really be any process.
+We can study the motion of a balloon on earth at 20 Celsius.
+This would not be possible on the surface of the sun as our balloon would not exist.
+The process has to be such that the system is recognizable to be the same.
+In our framework, we need some type of homomorphism between $\tdomain_t$ and $\tdomain_{t + \Delta t}$ that preserves narrowness and fineness, that tells us the system is the same.
+
+Second we want to make sure that the state space of the system really tells us only about the system itself.
+Within a specific process, for example, saying \statement{at time $t$, the velocity of object $A$ is $v$} may also mean that \statement{at time $t$, the velocity of object $B$ is $v$} because they are tied together.
+We say two systems are \textbf{independent} at a particular time if the description of one system does not constrain the description of the other.
+In terms of measure $\mu_{\stmt[u]_a \AND \stmt[u]_b}(x_a \AND x_b) = \mu_{\stmt[u]_a}(x_a) \mu_{\stmt[u]_b}(x_b)$ must factorize.
+If we want our state space to describe only the system at hand, then we are assuming that states are defined on the assumption that the system is independent at least from all other systems with the chosen processes.
+
+On the other side, a state of a system is never really the description of the whole system, but of the part that is relevant to the process at hand.
+If we study the motion of the balloon, position and velocity may be enough; if we study it as a thermodynamic system, pressure and volume may be enough.
+The idea is that the state captures the system at a particular level of granularity and the process does not depend, is not sensitive, to the finer dynamics.
+We say a system is in \textbf{equilibrium} if the state tells us nothing about the
+
+We stress that the independence cannot be, in general, independent throughout the whole process or there cannot be interactions between the systems.
+By the same token, the finer dynamics may play a part throughout the process.
+The constraint is only on the initial and final states, where the state space is being constructed.
+We also stress that these constraints make the state definition contingent upon the type of processes one studies.
+If we want to study a process where two systems are not independent, one that depends on the finer dynamics, our state definition will have to change to accommodate.
+
+ In all processes we are considering, then, the system must be indAny other system would be independent If we found other systems that were always correlated For example, if  There can't be other systems that participate in the process whose state
+
+We should caution that
+
+The simplest setting is the follow
+
+ into Our goal now is to construct state spaces for a system.
+
+ Consequently any statement that corresponds to finitely many evolutions (e.g.
+\statement{the full trajectories is either $y=t \cdot 1 m/s$ or $y=-t \cdot 1 m/s$}) will be comparable to all other statements.
+
+Also note that in our ideal gas example one of the quantities was determined with infinite precision.
+If we gave a range for all quantities, we would compare areas in that space, which we can always do.
+Statements with finite precision measurements are the only ones we can experimentally verify.
+So, we will require that all experimentally verifiable statements are comparable.
+We call \textbf{experimental domain} for the process the subset $\edomain[P] \subseteq \tdomain[P]$ of statements that can be, at least in line of principle, experimentally verified.\footnote{Yet, as we will see later, the precision of measurements will be quantified in terms of number of evolutions, not the range/uncertainty on the variables.}
+
+While we leave the full mathematical characterization in the mathematical section, there are a few things of note.
+Verifiable statements will correspond, in terms of sets of $E$, to finite and infinite volumes.
+Requiring those to be comparable means that $E$ is a measurable space and in phase space, for example, the standard measure will be recovered.
+As the process domain corresponds to a $\sigma$-algebra on $E$, the experimental domain corresponds to a topology on $E$.
+Therefore notions of continuity can be recovered as well.
+The process domain is the closure of the experimental domain under standard Boolean operations.
+This means that all the process domain is fully specify by, at least in principle, experimentally verifiable statements.
+The space does not contain any assertion that cannot ultimately be defined in terms of experiments.
+Lastly, requiring that, given enough time, each statement will be verified means the experimental domain must be generated by countably many statements.
+This means the topology is second countable and one is limited to work with ``nice'' spaces (e.g.
+spaces of continuous functions, separable Hilbert spaces, ...) not because they are ``nice'', but because they are the only physically meaningful ones.
+
+To sum up, we have two foundational object.
+The first object is the process domain $\tdomain[P]$ which is a Boolean algebra that contain all the possible description of our process.
+A subset $\edomain[P]$ contains those that are experimentally verifiable, which contain all the information in the full domain.
+The narrowest statements in $\tdomain[P]$ correspond to the evolutions $E$, that provide the full description at all times.
+The second object is the operator $\finer$ which gives us a way to compare two statements and decide which one gives a more refined description.
+Though two statements will not in general be comparable, all verifiable statements are comparable and all evolutions provide an equally refined description.
+
+This framework imposes very few requirements which are of very general applicability.
 It does not matter what specific quantities are used, or whether the process is from physics, chemistry, control systems, ecology, economics, sociology, or some other realm we have yet to discover.
+In the end, if we have a process we need to model, our theory will need at least to provide those objects.
+This means that the results we find will have broad applicability as well.
 
@@
 As the process extends over time, it is natural to want to organize the statements in our process domain into partial domains, one for each moment in time.
-Given our time parameter $t \in T \subseteq \mathbb{R}$, we can imagine to carve a domain $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
+Given our time parameter $t \in T \subseteq \mathbb{R}$, we can imagine to carve a subdomain $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
+Any logical operation on statements at fixed $t$ is also another statement at fixed $t$ therefore $\tdomain_t$ is a $\sigma$-complete Boolean sub-algebra of $\tdomain[P]$.
+The set $X_t \subset \tdomain_t$ of the narrowest statements will give the most precise description at that time: these are the possible states of our system.
+
+We should note that, technically, the statements are not defined at an infinitesimal moment $t$, rather within a finite interval $[t, t + \Delta t]$.
+This represent the time-scale at which the process will be described, meaning the description within $\tdomain_t$ should not be sensitive to what happens at faster scales.
+The quantities used to represent states at time $t$, then, should really be thought as averages within $\Delta t$.
+
+We should also note that there is nothing at this point requiring us that the set of descriptions $\tdomain_t$ and states $X_t$ will be the same at all times.
+
+What we want to do now is relate states at different times.
+In particular we want to formally define the notion of deterministic process: one in which, given the state at a given time, we can always predict the state at all future times.
+This can be done through the use of the narrowness operator: $\stmt_{t_1} \narrower \stmt_{t_2}$ means that all evolutions compatible with $\stmt_{t_1}$ are also compatible with $\stmt_{t_2}$.
+In other words, if we find $\stmt_{t_1}$ to be true we will also find $\stmt_{t_2}$ to be true as well.
+Therefore a process is deterministic if, for all $t_1, t_2 \in T$, given $x_{t_1} \in X_{t_1}$ we can find $x_{t_2} \in X_{t_2}$ such that $x_{t_1} \narrower x_{t_2}$.
 
@@
 
+NOTE: measure is invariant on Hamiltonian evolution, but not under irreversible evolution. Can't have the same measure at all times.
+
 \section{Math section}
@@
 
+Things I want to prove:
+
+\begin{prop}
+	Let $\tdomain[P]$ be a deterministic process over $\{X_t\}_{t \in T}$. Then $S(x(t))$ is a monotonically increasing function.
+\end{prop}
+
+\begin{prop}
+	Let $\tdomain[P]$ be a reversible process over $\{X_t\}_{t \in T}$. Then $S(x(t))$ is a monotonically decreasing function.
+\end{prop}
+
+\begin{prop}
+	Let $\tdomain[P]$ be a deterministic and reversible process over $\{X_t\}_{t \in T}$. Then $S(x(t))$ is constant.
+\end{prop}
+
+\begin{prop}
+	Let $\tdomain[P]$ be a deterministic process over $\{X_t\}_{t \in T}$. Then $S(x(t))$ is a monotonically increasing function.
+\end{prop}
+
 \bibliography{bibliography}

### 2020-06-16 Articles/2019-HamiltonianMechanicsIsConservationOfEntropy/arxiv/article.tex
Minor corrections

@@
 As vectors are coordinate invariant, $n$-forms are coordinate invariant as well so we can write $\mu(V)=\int_V \nu$ with no reference to coordinates.
-If $e_a$ are the basis vectors associated with $\xi^a$ and $e^b$ are linear functions such that $e^b(e_a)=\delta_a^b$, then we can express each $n$-form as $\rho(\xi^a)e_1\wedge e_2 \wedge ...
-\wedge e_n$.
+If $e_a$ are the basis vectors associated with $\xi^a$ and $e^b$ are linear functions such that $e^b(e_a)=\delta_a^b$, then we can express each $n$-form as $\rho(\xi^a)e^1\wedge e^2 \wedge ...
+\wedge e^n$.
 The density $\rho(\xi^a)$ is the component of the $n$-form expressed in the $\xi^a$ coordinates in a way analogous to components of a vector.\footnote{Technically, the form is linear so it changes with the Jacobian determinant and not its absolute value.
@@
 The density and information entropy must be invariant under arbitrary change of coordinates, not arbitrary change of state variables.
-The insight is that one choice of coordinate can influence more then one state variable.
+The insight is that one choice of coordinate can influence more than one state variable.
 But how many can it influence?
@@
 \hat{q} &= \hat{q}(q) \\
-\{\hat{q}, \hat{p}\} &= 1 = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} \\
+|J| = \{\hat{q}, \hat{p}\} &= 1 = \frac{\partial \hat{q}}{\partial q} \frac{\partial \hat{p}}{\partial p} \\
 \dfrac{\partial \hat{p}}{\partial p} &= \frac{\partial \hat{q}}{\partial q} ^{-1} = \frac{\partial q}{\partial \hat{q}} \\
@@
 That is, the particle state space $(\mathcal{S}, \omega)$ is a symplectic manifold.
-And, as any symplectic manifold, we can find Darboux variables $q^i, p_j$ and express the two-form as
+And, as any symplectic manifold, we can find Darboux variables $\xi^a = \{q^i, p_j\}$ and express the two-form as
 \begin{equation}

### 2020-06-16 Articles/2020-Refactoring/Source/refactoring-article.tex
Starting to rewrite article on general theory

--- 2019-04-02 Articles/TowardsAGeneralMathematicalTheoryOfExperimentalScience/article.tex
@@
-\newif\ifarchive
-\archivetrue
-%\archivefalse
+\documentclass[letterpaper]{article}
 
-\documentclass[%
-reprint,
-%superscriptaddress,
-%groupedaddress,
-%unsortedaddress,
-%runinaddress,
-%frontmatterverbose, 
-%preprint,
-%preprintnumbers,
-%nofootinbib,
-%nobibnotes,
-%bibnotes,
-amsmath,amssymb,
-aps,
-%pra,
-prx,
-%rmp,
-%prstab,
-%prstper,
-%floatfix,
-]{revtex4-1}
+\input{../../../include/basicstyle}
+\input{../../../include/theorems}
+\input{../../../include/logic}
 
-%% Language and font encodings
-%\usepackage[english]{babel}
-%\usepackage[utf8x]{inputenc}
-%\usepackage[T1]{fontenc}
-
-%% Sets page size and margins
-%\usepackage[letterpaper,top=1in,bottom=1in,left=1in,right=1in,marginparwidth=0.5in]{geometry}
-
-%\usepackage{lineno,hyperref}
 \usepackage{tikz}
-
-% Theorem definitions using amsthm
-\usepackage{amsmath, amsthm, amsfonts}
-
-\newenvironment{rationale}{\emph{Rationale}.}{\qed}
-\newenvironment{justification}{\emph{Justification}.}{\qed}
-\renewenvironment{proof}{\emph{Proof}.}{\qed}
-
-\theoremstyle{plain}% default 
-\newtheorem{thrm}{Theorem}[section] 
-\newtheorem{prop}[thrm]{Proposition} 
-\newtheorem{coro}[thrm]{Corollary}
-
-\theoremstyle{definition}
-\newtheorem{defn}[thrm]{Definition}
-\newtheorem{axiom}[thrm]{Axiom}
-\newtheorem*{principle*}{Principle of scientific objectivity}
-\theoremstyle{remark}
-\newtheorem*{rem}{Remark}
-
-\usepackage[only,llbracket, rrbracket,llparenthesis,rrparenthesis]{stmaryrd}
 
@@
 
-% LOGIC symbols
-% -------------
-
-% Allows to create negation symbols
-\let\vec\undefined
-\usepackage{MnSymbol}
-
-\DeclareMathOperator{\truth}{truth}
-\DeclareMathOperator{\possFn}{poss}
-\DeclareMathOperator{\result}{result}
-\DeclareMathOperator{\idFn}{id}
-
-\def\Bool{\mathbb{B}}
-\def\TRUE{\textsc{true}}
-\def\FALSE{\textsc{false}}
-
-\def\SUCCESS{\textsc{success}}
-\def\FAILURE{\textsc{failure}}
-\def\UNDEF{\textsc{undefined}}
-
-% Symbols for statements set
-\def\stmtSet{\mathcal{S}}
-\def\vstmtSet{\mathcal{S}_\textsf{v}}
-\def\dstmtSet{\mathcal{S}_\textsf{d}}
-
-% Symbols for tautology and contradiction
-\def\tautology{\top}
-\def\contradiction{\bot}
-
-% Symbols for "compatibility" and "incompatibility"
-\def\comp{\doublefrown}
-\def\ncomp{\ndoublefrown}
-
-% Symbols for "narrower" and "wider"
-\def\narrower{\preccurlyeq}
-\def\nnarrower{\npreccurlyeq}
-\def\broader{\succcurlyeq}
-\def\nbroader{\nsucccurlyeq}
-
-% Symbol for "independent" and "correlated"
-\def\indep{\upmodels}
-\def\nindep{\nupmodels}
-
-% Aliases for logical operations
-\def\AND{\wedge}
-\def\bigAND{\bigwedge}
-\def\OR{\vee}
-\def\bigOR{\bigvee}
-\def\NOT{\neg}
-
-% Formatting for statements
-\newcommand{\stmt}[1][s] {\mathsf{#1}}
-% Formatting for experimental tests
-\newcommand{\expt}[1][e] {\mathsf{#1}}
-\newcommand{\exptSet}{\mathcal{E}}
-% Formatting for observations
-\newcommand{\obs}[1][s] {\mathsf{#1}}
-
-% Formatting for possibilities
-\def\logCtx{\mathcal{S}}
-\newcommand{\pAss}[1][\mathcal{S}] {\mathcal{A}_{#1}}
-
-% Formatting for experimental domain
-\newcommand{\edomain}[1][D] {\mathcal{#1}}
-
-% Formatting for theoretical domain
-\newcommand{\tdomain}[1][D] {\bar{\mathcal{#1}}}
-
-\newcommand{\basis}[1][B] {\mathcal{#1}} % Basis
-
-% Formatting for experimental relationships
-\newcommand{\erel}[1][r] {#1}
-
-% Formatting for sentence statements
-\newcommand{\statement}[1] {\emph{``#1''}}
-
 \begin{document}
 
-\title{Towards a general mathematical theory of experimental science}
+\title{A Refactoring - new attempt}
 
 \author{Gabriele Carcassi}
-\email{carcassi@umich.edu}
+%\email{carcassi@umich.edu}
 \author{Christine A. Aidala}
-\affiliation{
-	Physics Department \\
-	University of Michigan \\
-	Ann Arbor, MI 48109 \\
-}
+%\affiliation{
+%	Physics Department \\
+%	University of Michigan \\
+%	Ann Arbor, MI 48109 \\
+%}
 
@@
 
-\section{The principle of scientific objectivity}
+\section{A standard for precision?}
 
-As the starting point of our general theory, we introduce the following guiding principle:
+Mathematics has evolved a set of rules and standards for what constitutes a well-posed mathematical theory.
+There is no such standard in science in general and in physics in particular.
+In fact, given that some fields are mostly driven by data and heuristics while others are driven by mathematical formalism, we have found divergent opinions in what can and should be formalized.
+We therefore had to develop our own standard, which is based upon the following insights.
 
-\begin{principle*}
-Science is universal, non-contradictory and evidence based.
-\end{principle*}
+\textit{Formal systems cannot fully capture scientific theories.} The same mathematical structure (e.g.
+the equation $a= bc$) can refer to multiple physical relationships (e.g.
+Newton's second law or Ohm's law) which cannot be reconstructed solely on the basis of the formal relationships.
+Moreover, qualitative reasoning in many fields can be carried out independently.
+In this light, formal systems are a poor starting point for a scientific theory.
 
-This means that science concerns itself only to the study of assertions that have a well defined truth value which can be verified experimentally.
-The issue at hand is to formally capture this informal intuition.
-We begin with the following common definition.
+\textit{Formal systems in science should be justified, not interpreted.} We want to start from the informal definitions of the physical objects, be clear as to what aspects are formalized and why that formalization follows.
+The formal system, then, is a tool useful to check for consistency and clarification of particular aspects of the theory.
+It is useful only insofar its connection to the physical content is clear.
 
-\begin{defn}
-	The \textbf{Boolean domain} is the set $\mathbb{B} = \{\FALSE, \TRUE\}$ of all possible truth values.
-\end{defn}
+Our framework will therefore consist of both informal and formal parts.
+We distinguish two ways new concepts are introduced: axioms and definitions.
+\textbf{Axioms} introduce new elements: they define new concepts and what aspects are captured in the formal system.
+\textbf{Definitions} further characterize already introduced objects, without introducing new elements in the formal system.
+Both axioms and definitions will consist of two parts.
+The first gives the intuitive, informal description while the second gives the formal characterization.
+The formal part will necessarily not capture the entirety of the informal concept.
+Where needed, a \textbf{justification} will accompany axioms and definitions.
+These are informal arguments that explain why the formal objects must have those formal properties.
+All \textbf{propositions} and \textbf{theorems} are declared and proven in the formal system, and they deal with objects that are now both mathematically and scientifically well specified.
 
-Next we need to define our truth bearer.
-In mathematical logic, this is typically a well formed formula.
+The overall goal is to keep the axioms to a minimum and to a level of simplicity that is easily checked. This minimizes the part of the framework that can potentially be open to inconsistencies.
+
+Axioms, definitions and justifications are included in the mathematical section, with the rest of the formal elements. They should be seen as integral part of the framework.
+
+\section{Basic formal structures?}
+
+The starting point of our framework is the \textbf{principle of scientific objectivity} which states \statement{Science is universal, non-contradictory and evidence based}.
+Every physical theory, then, must give us a set of logically sound statements linked with some notion of experimental verification.
+The content of the statements, what can be verified and how, will vary from theory to theory, but logical consistency and verifiability must be common to all of them.
+Therefore we will have two set of axioms: the \textbf{logical consistency axioms} and the \textbf{experimental verifiability axioms}.
+We start with the first set.
+
+\subsection{Logical consistency}
+
+The \textbf{axiom of context} establishes that \textbf{statements} are organized into \textbf{logical contexts}.
+It is the context that defines their meaning and logical relationships.
+The definition of a physical object, for example, may change depending on our goal: if we measure the mass of the electron, its value is assumed to be unknown; if we perform particle identification in a detector, the mass of the electron is assumed to be a particular value.
+Each context has to be such that the meaning of each statement has no ambiguity and therefore it is associated with a single, well-defined truth value.
+This is, in general, not known: it cannot be proven or deduced.
+Ultimately, it is the role of experimentation to find it.
+
+The \textbf{axiom of possibility} establishes that not all truth assignments are consistent with the meaning of the statements.
+For example, it is not possible for \statement{the particle is an electron} and \statement{the particle is a positron} to be true at the same time.
+As the context must fully specify the meaning of each statement, all these relationships must be well-specified.
+A context therefore will define the \textbf{possible assignments} for all statements, and the actual truth must be one of them.
+
+Lastly, the \textbf{axiom of closure} establishes that we can always find a statement whose truth depends on the truth of others through a chosen Boolean function.
+For example, we can find the negation of a statement, the conjunction (logical AND) of any two statements, and so on.
+
+These three axioms introduce the basis of a logic system all scientific theories must adhere to.
+We can visualize a logical context as a table.
+Suppose we have a context $\logCtx$ which contains at least the following statements:
+\begin{description}
+	\item $\stmt_1=$\statement{that animal is an animal}
+	\item $\stmt_2=$\statement{that animal is a cat}
+	\item $\stmt_3=$\statement{that animal is a mammal}
+	\item $\stmt_4=$\statement{that animal has three bones in the middle ear}
+	\item $\stmt_5=$\statement{that animal is a dog}
+	\item $\stmt_6=$\statement{that animal is black}
+	\item $\stmt_7=$\statement{that animal is a plant}
+\end{description}
+We would have:
+\begin{center}
+	\begin{tabular}{c|c|c|c|c|c|c|c}
+		$\stmt_1$ & $\stmt_2$ & $\stmt_3$ & $\stmt_4$ & $\stmt_5$ & $\stmt_6$ & $\stmt_7$ & ... \\
+		\hline
+		T & T & T & T & F & T & F & ... \\
+		T & T & T & T & F & F & F & ... \\
+		T & F & T & T & F & T & F & ... \\
+		T & F & F & F & T & T & F & ... \\
+		T & F & F & F & F & F & F & ... \\
+		... & ... & ... & ... & ... & ... & ... \\
+	\end{tabular}
+\end{center}
+where each line represents a possible assignment, each column a statement and each cell the truth value for each statement under the given assignment.
+
+We can give some further definitions, which we introduce with examples based on the previous table.
+We say $\stmt_3$ and $\stmt_4$ are \textbf{equivalent}, noted $\stmt_3 \equiv \stmt_4$ because they have the same truth value in each assignment (only mammals have three bones in the middle ear).
+We say $\stmt_2$ is \textbf{narrower} than $\stmt_3$ (i.e.
+more specific), noted $\stmt_2 \narrower \stmt_3$, because whether the first one is true the second one is true as well (if the animal is a cat, it is also a mammal).
+We say $\stmt_2$ and $\stmt_5$ are not \textbf{compatible}, noted $\stmt_2 \ncomp \stmt_5$, because there is no assignment in which they are both true (if the animal is a cat, it is not a dog).
+We say $\stmt_2$ is independent from $\stmt_6$, noted $\stmt_2 \indep \stmt_6$, because the truth value of one does not constrain the truth value of the other (knowing whether the animal is a cat or not does not tell us whether it is black or not).
+We say $\stmt_1$ is a \textbf{tautology} $\tautology$, noted $\stmt_1 \equiv \tautology$, because it is true in every assignment.
+We say $\stmt_7$ is a \textbf{contradiction} $\contradiction$, noted $\stmt_7 \equiv \contradiction$, because it is false in every assignment.
+All others are \textbf{contingent}.
+
+From now on, we will not differentiate between two statements that are equivalent.
+That is, we will implicitly refer to the equivalence classes.
+With that in mind, one can show that a logical context is a \textbf{complete Boolean algebra} and that $\narrower$ is a partial order, the one corresponding to the lattice of the algebra.
+While this should not be surprising, the point is that we start from premises that are easily justified and that give a clear understanding of what the formal object represents.
+
+This will be essential when adding new definitions.
+Any other formal structure imposed on the statements, in fact, is capturing other properties of the same objects and will have to ``play nice'' with the above logical structure.
+This means that the notion of equivalence will, in practice, become an isomorphism within the new structure; the ordering imposed by numeric quantities will be a specialized version of the one imposed by $\narrower$; the notion of independence will relate to the notion of independence in statistic, probability, linear algebra, dynamical system, and so on.
+Understanding this simple logical structure, then, means getting insights in many different areas and in how they are ultimately connected.
+
+% In mathematical logic, this is typically a well formed formula.
 This cannot work for us: what we are interested in is the meaning of the assertion and not how it is expressed.
@@
 
-\begin{axiom}\label{ax_statement}
-	A \textbf{statement} $\stmt$ is an assertion that is either true or false.
-A \textbf{logical context} $\logCtx$ is a collection of statements with well defined logical relationships.
-Formally, a logical context $\logCtx$ is a collection of elements called statements upon which is defined a function $\truth: \logCtx \to \Bool$.
-\end{axiom}
-
-Note how the first part of the definition captures the informal meaning of what we are describing, while the second part captures the part that is formalized.
+%Note how the first part of the definition captures the informal meaning of what we are describing, while the second part captures the part that is formalized.
 This pattern will be present in most of our definitions and it serves to clarify both what is being formalized and how.
@@
 
-While in math the truth value is the focus of the logic system, in science it is generally established experimentally.
-But our scientific model may constrain certain statements or statement combinations to be ruled out.
-For example, the statements \statement{this animal is a dog} and \statement{this animal is a cat} can both be either true or false, but they can't both be true.
-That is, the role of our logic system is not to keep track of what is true and false, but of what cannot possibly be true and cannot possibly be false.
-We need to keep track of these relationships so that we can never have inconsistencies or paradoxes.
-
-\begin{axiom}\label{ax_possible_assignemetns}
-	A \textbf{possible assignment} for a logical context $\logCtx$ is a map $a : \logCtx \to \Bool$ that assigns a truth value to each statement in a way consistent with the content of the statements.
-Formally, each logical context comes equipped with a set $\pAss \subseteq \Bool^{\logCtx}$ such that $\truth \in \pAss$.
-A map $a : \logCtx \to \Bool$ is a possible assignment for $\logCtx$ if $a \in \pAss$.
-\end{axiom}
-
-Note that the possible assignments are just hypotheticals.
+%Note that the possible assignments are just hypotheticals.
 They are not actual physical entities.
@@
 
-\begin{defn}
-	A \textbf{tautology} $\tautology$ is a statement that is true in every possible assignment.
-\end{defn}
-
-\begin{defn}
-	A \textbf{contradiction} $\contradiction$ is a statement that is false in every possible assignment.
-\end{defn}
-
-We also need to express relationships between the truth of different statements.
-For example, if we assign true to \statement{this animal is a dog} then we cannot assign true to \statement{this animal is not a dog}.
-Therefore we introduce the following:
-
-\begin{axiom}\label{ax_functions_of_statement}
-	We can always construct a statement whose truth value arbitrarily depends on an arbitrary set of statements.
-Formally, let $S \subseteq \logCtx$ be a set of statements and $f_{\Bool} : \Bool^S \to \Bool$ an arbitrary function from an assignment of $S$ to a truth value.
-Then we can always find a statement $\bar{\stmt} \in \logCtx$ such that
-	$$a(\bar{\stmt}) = f_{\Bool}(\{a(\stmt)\}_{\stmt \in S})$$
-	for every possible assignment $a \in \pAss$.
-\end{axiom}
-
-We will use the standard symbols $\NOT$, $\AND$, $\OR$ to indicate the negation (logical NOT), conjunction (logical AND) and disjunction (logical OR).
-With these three axioms, we can rederive all of the rules of classical logic.
-First we define our equivalence.
-
-\begin{defn}
-	Two statements $\stmt_1$ and $\stmt_2$ are \textbf{equivalent} $\stmt_1 \equiv \stmt_2$ if they must be equally true or false simply because of their content.
-Formally, $\stmt_1 \equiv \stmt_2$ if and only if $a(\stmt_1) = a(\stmt_2)$ for all $a \in \pAss$.
-\end{defn}
-
-From this definition one can prove the following two propositions.
-
-\begin{prop}
-	Statement equivalence satisfies the following properties:
-	\begin{itemize}
-		\item reflexivity: $\stmt \equiv \stmt$
-		\item symmetry: if $\stmt_1 \equiv \stmt_2$ then $\stmt_2 \equiv \stmt_1$
-		\item transitivity: if $\stmt_1 \equiv \stmt_2$ and $\stmt_2 \equiv \stmt_3$ then $\stmt_1 \equiv \stmt_3$
-	\end{itemize}
-	and is therefore an \textbf{equivalence relationship}.
-\end{prop}
-
-\begin{prop}\label{boolean_properties}
-	The set of all statements $\mathcal{S}$ satisfies the following properties:
-	\begin{itemize}
-		\item associativity: $a \OR (b \OR c) \equiv (a \OR b) \OR c$, $a \AND (b \AND c) \equiv (a \AND b) \AND c$
-		\item commutativity: $a \OR b \equiv b \OR a$, $a \AND b \equiv b \AND a$
-		\item absorption: $a \OR (a \AND b) \equiv a$, $a \AND (a \OR b) \equiv a$
-		\item identity: $a \OR \contradiction \equiv a
-		$, $a \AND \tautology \equiv a$
-		\item distributivity: $a \OR (b \AND c) \equiv (a \OR b) \AND (a \OR c)$, $a \AND (b \OR c) \equiv (a \AND b) \OR (a \AND c)$
-		\item complements: $a \OR \NOT a \equiv \tautology$, $a \AND \NOT a \equiv \contradiction$
-		\item De Morgan: $\NOT a \OR \NOT b \equiv \NOT (a \AND b)$, $\NOT a \AND \NOT b \equiv \NOT (a \OR b)$
-	\end{itemize}
-	This, by definition, means $\mathcal{S}$ is a \textbf{Boolean algebra}.
-\end{prop}
-
-One can also prove that the Boolean algebra is complete, and therefore it generalizes to the infinite case.
-
-This system not only allows us to use formal classical logic while keeping the statements informal, but is also equipped to capture causal relationships.\footnote{Note that the axioms prevent any form of paradox, simply because a possible assignment must exist.
+%This system not only allows us to use formal classical logic while keeping the statements informal, but is also equipped to capture causal relationships.\footnote{Note that the axioms prevent any form of paradox, simply because a possible assignment must exist.
 Therefore situations like \statement{this statement is not true} are ruled out simply because they cannot satisfy the axioms.} For example, $\stmt_1=$\statement{the thermometer indicator is between 24 C and 25 C} and $\stmt_2=$\statement{the temperature is between 24 C and 25 C} are equivalent because one is true if and only if the other is true (on the assumption that our thermometer is actually working).
@@
 
-\begin{defn}
-	Given two statements $\stmt_1$ and $\stmt_2$, we say that:
-	\begin{itemize}
-		\item $\stmt_1$ \textbf{is narrower than} $\stmt_2$ (noted $\stmt_1 \narrower \stmt_2$) if $\stmt_2$ is true whenever $\stmt_1$ is true.
-		\item $\stmt_1$ \textbf{is broader than} $\stmt_2$ (noted $\stmt_1 \broader \stmt_2$) if $\stmt_2 \narrower \stmt_1$.
-		\item $\stmt_1$ \textbf{is compatible to} $\stmt_2$ (noted $\stmt_1 \comp \stmt_2$) if there is an assignment in which they are both true.
+\subsection{Experimental verifiability}
 
-	\end{itemize}
-	The negation of these properties will be noted by $\nnarrower$, $\nbroader$ , $\ncomp$ respectively.
-\end{defn}
-\begin{defn}
-	The elements of a set of statements $S \subseteq \logCtx$ are said to be \textbf{independent} (noted $\stmt_1 \indep \stmt_2$ for a set of two) if the assignment of any subset of statements does not depend on the assignment of the others.
-That is, a set of statements $S \subseteq \logCtx$ is independent if given a family  $\{t_{\stmt}\}_{\stmt \in S}$ such that each $t_{\stmt} \in \Bool$ is a possible assignment for the respective $\stmt$ we can always find $a \in \pAss$ such that $a(\stmt) = t_{\stmt}$ for all $\stmt \in S$.
-\end{defn}
+Having seen the logical consistency axioms, we turn our attention to the experimental verifiability axioms.
 
-For example, given the following statements:
-\begin{description}
-	\item $\stmt_1=$\statement{that animal is a cat}
-	\item $\stmt_2=$\statement{that animal is a mammal}
-	\item $\stmt_3=$\statement{that animal is a dog}
-	\item $\stmt_4=$\statement{that animal is black}
-\end{description}
-we have the $\stmt_1$ $\narrower$ $\stmt_2$ (i.e.
-it is more specific), $\stmt_1$ $\ncomp$ $\stmt_3$ (i.e.
-they cannot be true at the same time) and $\stmt_1$ $\indep$ $\stmt_4$ (i.e.
-the truth of one tells us nothing about the truth of the other).
-
-Another interesting result is that statement narrowness imposes a partial order on the set of all statements.
-
-\begin{prop}
-	Statement narrowness satisfies the following properties:
-	\begin{itemize}
-		\item reflexivity: $s \narrower s$
-		\item antisymmetry: if $s_1 \narrower s_2$ and  $s_2 \narrower s_1$ then $s_1 \equiv s_2$
-		\item transitivity: if $s_1 \narrower s_2$ and $s_2 \narrower s_3$ then $s_1 \narrower s_3$
-	\end{itemize}
-	and is therefore a \textbf{partial order}.
-\end{prop}
-
-As every element in our general theory will be constructed upon statements, these operations are important as they will characterize all those constructions.
-For example, if we quantify the precision of a set of statements, it will need to be ordered in a way that is compatible with narrowness.
-If we define statistical independence, it will have to be defined in a way that is compatible with statement independence.
-In the same way that set theory defines concepts common across all mathematics, the general theory defines basic concepts that are common to all scientific theories.
-
-\section{Verifiable statements and experimental domains}
-
-We have the tools for dealing with assertions that are universal and non-contradictory, now we have to develop the tools for those that are evidence based as well.
-
-\begin{axiom}\label{ax_verifiable_statements}
-	A \textbf{verifiable statement} is a statement that can be shown to be true experimentally.
-Formally, a statement $\stmt$ is verifiable if it is part of the subset $\stmt \in \vstmtSet \subset \stmtSet$ of all verifiable statements.
-\end{axiom}
-
-Physically, this means that we have a repeatable procedure that anybody can execute and that always gives the same result.
-If the statement is true, then this experimental test must terminate successfully.
-Note that, in general, the test may not terminate if the statement is false.
-Consider the following:
+The \textbf{axiom of verifiability} establishes that some statement are experimentally \textbf{verifiable}, meaning we have a test at our disposal that, if the statement is true, will terminate successfully in finite time.
+If the statement is not true, the test may terminate unsuccessfully or never terminate.
+To give an example, consider the following:
 \begin{enumerate}
@@
 absence of evidence is not evidence of absence).
-Because of non-termination, failure to verify is not verification of the negation.
+The axiom also states that tautologies and contradictions are verifiables (i.e.
+they can be associated with trivial tests) and that if one statement is equivalent to a verifiable one, then it is also verifiable (i.e.
+we can use the test of the second since either both statements are true or both are false).
+
+Termination in finite time is important as it imposes a different logical structure.\footnote{This is broadly recognized in computer science theory and in constructive logic.} Because of non-termination, failure to verify is not verification of the negation.
 That is, the negation of a verifiable statement is not necessarily a verifiable statement.
 
-Given two statements, though, we can verify their conjunction simply by verifying both: if they are both true, both their tests will terminate and verify the conjunction.
-But we cannot extend this to an infinite number of statements as we would never terminate.
-We can also verify the disjunction of two statements: once one test terminates we are done.
-And because we only need one test to terminate, we can generalize to a countable number of verifiable statements by following this procedure:
-\begin{enumerate}
-	\item initialize $n$ to 1
-	\item for each $i=1..n$
-	\begin{enumerate}
-		\item run the test for $\stmt_i$ for $n$ seconds
-		\item if it terminates successfully then terminate successfully
-	\end{enumerate}
-	\item increment $n$ and go to step 2
-\end{enumerate}
-This procedure will run all tests for an arbitrary amount of time.
-Therefore, if one statement is true, it will make the test terminate successfully, even if all other tests would never terminate.
-In light of this, we can set the following axioms.
+The \textbf{axiom of finite conjunction verifiability} establishes that, given a finite set of statements, we can verify their conjunction (i.e.
+the logical AND).
+In fact, we can execute one test at a time, and if they are all successful, they will be so in finite time therefore the conjunction will be verified as well.
+This will not extend to the infinite case.
 
-\begin{axiom}\label{ax_verifiable_AND}
-	The conjunction of a finite collection of verifiable statements is a verifiable statement.
-Formally, let $\{\stmt_i\}_{i=1}^{n} \subseteq \vstmtSet$ be a finite collection of verifiable statements.
-Then the conjunction $\bigAND\limits_{i=1}^{n} \stmt_i \in \vstmtSet$ is a verifiable statement.
-\end{axiom}
-	\begin{axiom}\label{ax_verifiable_OR}
-	The disjunction of a countable collection of verifiable statements is a verifiable statement.
-Formally, let $\{\stmt_i\}_{i=1}^{\infty} \subseteq \vstmtSet$ be a countable collection of verifiable statements.
-Then the disjunction $\bigOR\limits_{i=1}^{\infty} \stmt_i \in \vstmtSet$ is a verifiable statement.
-\end{axiom}
+The \textbf{axiom of countable disjunction verifiability} establishes that, given a countable set of statements, we can verify their disjunction (i.e.
+the logical OR).
+In fact, if the disjunction is true then one statement must be true.
+Once that is verified, we can terminate regardless of how many other statements are there in the disjunction.
+The set of statements must be countable, though, or it may take infinite time to find the statement that is true.
 
-We could also define decidable statements as those for which falsehood can also be tested experimentally. Table \ref{tab:algebras} compares the different algebras.
+On top of these axioms we can define \textbf{decidable statements} as the ones that can be both verified and falsified experimentally (i.e.
+their negation is a verifiable statement).
+One can prove that the finite disjunction and the finite conjunction of decidable statements is also decidable.
 
-\begin{table*}
-	\centering
-	\begin{tabular}{p{0.14\textwidth} p{0.08\textwidth} p{0.13\textwidth} p{0.22\textwidth} p{0.23\textwidth}}
-		Operator & Gate & Statement & Verifiable Statement & Decidable Statement  \\ 
-		\hline 
-		Negation & NOT & allowed & disallowed & allowed \\ 
-		Conjunction & AND & arbitrary  & finite & finite \\ 
-		Disjunction & OR & arbitrary  & countable & finite \\ 
-	\end{tabular}
-	\caption{Comparing algebras of statements.}\label{tab:algebras}
-\end{table*}
+We therefore have three types of algebras embedded one into the other.
+The logical context $\logCtx$ is a complete Boolean algebra; the set of all verifiable statement $\vstmtSet \subset \logCtx$ is, in general, not a Boolean algebra but a complete Heyting algebra; the set of all decidable statements $\dstmtSet \subset \vstmtSet \subset \logCtx$ is a (finite) Boolean algebra.
+The interplay between these structures is important and leads to topologies and $\sigma$-algebras, the foundations for geometry, Lie algebras, measure theory, probability theory and most of the mathematical structures widely used in physics.
 
-Now that we can verify statements one by one, we need to define what it means to verify a group of them.
-In principle, given a set of verifiable statements we can simply start testing them one after the other.
-However, if we are given a set of uncountable statements then, even if we have an indefinitely long time at our disposal, we will not be able to create a procedure that eventually reaches all statements.
-However, we may not need to actually run all tests for all statements.
-For example, if we found that $\stmt_1$ is true then there is no need to test any of its disjunctions like $\stmt_1 \OR \stmt_2$.
+\subsection{Domains and possibilities?}
 
-\begin{defn}
-	Given a set $\edomain$ of verifiable statements, $\basis \subseteq \edomain$ is a \textbf{basis} if the truth values of $\basis$ are enough to deduce the truth values of the set.
-Formally, all elements of $\edomain$ can be generated from $\basis$ using finite conjunction and countable disjunction.
-\end{defn}
+Finite time termination not only imposes a different logical structure, but it also limits the number of statements that can be verified.
+Even when given an arbitrarily long time, a countable set of statements is the most that can be verified.
+However, due to the logical relationship between statements, one does not need to test all of them: if \statement{the mass of the photon is less than $10^{-6}$ eV} is experimentally verified, there is no need to test whether \statement{the mass of the photon is greater than $10^{-6}$ eV}.
 
-Therefore it is the size of the basis that matters and not the size of the set.
-If the basis is countable we can keep going and, if any statement is true, it will eventually be verified experimentally.\footnote{We assume we are given an indefinitely long time because this handles limits and because we have not given constraints for how long or short a successful test can be (i.e.
-if we ran a test for three days without terminating, it may still successfully complete in three days and one minute).}
+Given a set $\edomain$ of verifiable statements, we call a \textbf{basis} $\basis \subset \edomain$ a subset whose tests are enough to verify the whole set.
+Formally, $\edomain$ can be generated from $\basis$ using finite conjunction and countable disjunction.
+We say $\edomain$ is an \textbf{experimental domain} if it is a maximal set (i.e.
+includes all the verifiable statements) generated from a countable basis.
 
-\begin{defn}
-	An \textbf{experimental domain} $\edomain$ represents all the experimental evidence that can be acquired about a scientific subject in an indefinite amount of time.
-Formally, it is a set of statements, closed under finite conjunction and countable disjunction, that includes precisely the tautology, the contradiction, and a set of verifiable statements that can be generated from a countable basis.
-\end{defn}
+An experimental domain represents all experimental evidence that can be acquired in an indefinite amount of time.
+It represents the most fundamental object in our framework and, if our scientific theory is evidence based, it must be fully specified by an experimental domain.
 
-These axioms and definitions formally characterize what we mean by evidence based.
-A scientific theory will be fully defined by a countable set of verifiable statements, the basis for an experimental domain.
+Not all statements in a scientific theory, though, are verifiable statement.
+For example, \statement{there is no extra-terrestrial life} is not something we can verified (at best we can only conclude we have found none where we looked), and neither is \statement{the mass of the photon is exactly 0 eV} (at best we can exclude is above a certain threshold).
+In general, while negations are not necessarily verifiable, we can still logically talk about them and they must be included in our theory.
 
-\section{Theoretical domains and possibilities}
+Given an experimental domain $\edomain$ we define its \textbf{theoretical domain} $\tdomain$ which includes all statements generated by $\edomain$ allowing negations as well.
+Theoretical domains are $\sigma$-complete Boolean algebras as they are closed under negation and countable conjunction and disjunction.
 
-While verifiable statements define what can be scientifically studied, not all interesting scientific statements are directly verifiable.
-Consider the two statements \statement{there exists extra-terrestrial life} and \statement{there is no extra-terrestrial life}.
-We can verify the first if we happen to find signs of life somewhere, but experimentally verifying the second is practically impossible.
-Yet, the second is still meaningful as a prediction: it predicts that the test for the first will never terminate.
-That is, while negations are not verifiable we can still logically talk about them when, for example, constructing truth assignments.
+For each theoretical statement $\tstmt \in \tdomain$, we can define its \textbf{verifiable part} $\ver(\tstmt)$, \textbf{falsifiable part} $\fal(\tstmt)$ and its \textbf{undecidable part} $\und(\tstmt)$.
+This allows us to precisely characterize the optimal testing strategy in terms of termination.
+For example, for a decidable statement like $\stmt=$\statement{there are three apples on the table}, the verifiable part will be equal to the statement itself $\ver(\stmt) \equiv \stmt$, the falsifiable part will be equal to its negation cases $\fal(\stmt) \equiv \NOT\stmt$ and the undecidable part will be the contradiction $\und(\stmt) \equiv \contradiction$.
+The experimental test for $\stmt$ always terminates either successfully or unsuccessfully.
+Instead, if $\stmt=$\statement{the mass of the photon is less than $10^{-17}$ eV}, we still have $\ver(\stmt) \equiv \stmt$ because it is a verifiable statement, but we will have $\fal(\stmt) \equiv$\statement{the mass of the photon is more than $10^{-17}$ eV} while $\und(\stmt) \equiv$\statement{the mass of the photon is exactly $10^{-17}$ eV}.
+The issue is that, due to finite precision, if the mass of the photon where exactly $10^{-17}$ eV, our confidence interval would only include the region below that number, so we would never be able to exclude it.
+So, the test would not terminate in that particular case.
+Lastly, if $\stmt=$\statement{the mass of the photon is a rational number as expressed in eV}, we have $\ver(\stmt) \equiv \contradiction \equiv \fal(\stmt)$ and $\und(\stmt) \equiv \tautology$.
+The only way to know whether the statement is true or not would be to know the exact value of the mass, which cannot be achieved in finite time.
+Therefore the test for $\stmt$ will never terminate.
+We call such statements \textbf{undecidable}.
 
-\begin{defn}
-	The \textbf{theoretical domain} $\tdomain$ of an experimental domain $\edomain$ is the set of statements that we can use to state predictions, which is constructed from $\edomain$ by allowing negation.
-We call \textbf{theoretical statement} a statement that is part of a theoretical domain.
-More formally, $\tdomain$ is the set of all statements generated from $\edomain$ using negation, finite conjunction and countable disjunction.
-\end{defn}
+The fact that a scientific theory can contain undecidable statements, statements about which nothing can be said experimentally, may seem counter-intuitive.
+Theoretical statements include all those statement for which, in principle, a test can be devised, regardless of whether the test actually terminates: the $\sigma$-completeness guarantees that.
+Note that if we closed under arbitrary union, not just countable, we would bring in statements for which a test cannot even be, in principle, constructed.
+These would be unphysical objects, and are excluded by our framework.\footnote{These would correspond to non-Borel sets.}
 
@@
 	Let $\edomain$ be an experimental domain and $X$ its possibilities.
-We define the map $U : \edomain \rightarrow 2^X$ that for each statement $\obs \in \edomain$ returns the set of possibilities compatible with it.
-That is, $U(\obs)\equiv\{ x \in X \, | \, x \comp \obs\}$.
-We call $U(\obs)$ the \textbf{verifiable set} of possibilities associated with $\obs$.
+We define the map $U : \edomain \rightarrow 2^X$ that for each statement $\stmt \in \edomain$ returns the set of possibilities compatible with it.
+That is, $U(\stmt)\equiv\{ x \in X \, | \, x \comp \stmt\}$.
+We call $U(\stmt)$ the \textbf{verifiable set} of possibilities associated with $\stmt$.
 \end{defn}
@@
 \begin{prop}
-	A statement $\obs \in \edomain$ is the disjunction of the possibilities in its verifiable set $U(\obs)$. That is, $\obs=\bigOR\limits_{x \in U(\obs)} x$.
+	A statement $\stmt \in \edomain$ is the disjunction of the possibilities in its verifiable set $U(\stmt)$. That is, $\stmt=\bigOR\limits_{x \in U(\stmt)} x$.
 \end{prop}

### 2020-06-21 Articles/2020-Refactoring/Source/refactoring-article.tex
More work on the article

@@
-\documentclass[letterpaper]{article}
+\documentclass[twocolumn]{article}
+\usepackage{authblk}
 
@@
 
-\author{Gabriele Carcassi}
 %\email{carcassi@umich.edu}
-\author{Christine A. Aidala}
-%\affiliation{
-%	Physics Department \\
-%	University of Michigan \\
-%	Ann Arbor, MI 48109 \\
-%}
+\author[1]{Gabriele Carcassi}
+\author[1]{Christine A. Aidala}
+\affil[1]{
+	Physics Department\\
+	University of Michigan \\
+	Ann Arbor, MI 48109
+}
 
 \date{\today}
+
+\maketitle
 
@@
 \end{abstract}
-
-\maketitle
 
@@
 
-On top of these axioms we can define \textbf{decidable statements} as the ones that can be both verified and falsified experimentally (i.e.
+On top of these axioms we can define \textbf{falsifiable statements} as the ones that can be disconfirmed experimentally (e.g.
+their negation is verifiable).
+We can also also define \textbf{decidable statements} as the ones that can be both verified and falsified experimentally (i.e.
 their negation is a verifiable statement).
@@
 The fact that a scientific theory can contain undecidable statements, statements about which nothing can be said experimentally, may seem counter-intuitive.
-Theoretical statements include all those statement for which, in principle, a test can be devised, regardless of whether the test actually terminates: the $\sigma$-completeness guarantees that.
+The issues is that theoretical statements include all those statement for which, in principle, a test can be devised, regardless of whether the test actually terminates: the $\sigma$-completeness guarantees that.
 Note that if we closed under arbitrary union, not just countable, we would bring in statements for which a test cannot even be, in principle, constructed.
-These would be unphysical objects, and are excluded by our framework.\footnote{These would correspond to non-Borel sets.}
+These would be truly unphysical objects, and are excluded by our framework.\footnote{These would correspond to non-Borel sets.}
 
-Note that the new statements provide no new information.
-In fact, all statements in the theoretical domain $\tdomain$ can be generated by negation, countable conjunction and countable disjunction from a basis $\basis$ of $\edomain$.
-As they provide all possible predictions, we focus on the ones that completely specify the truth value for all theoretical statements.
-For example, once we know that \statement{this animal is a cat} we know that \statement{this animal has whiskers}, that \statement{this animal has no feathers} and so on.
+As noted before, some statements are narrower than other, they provide a more refined description.
+We call \textbf{possibilities} the statements that give the most refined description without being contradictions, without being degenerate.
+If we knew one of the possibilities (e.g.
+\statement{this animal is a cat}) to be true, then we would know the truth value of all other statements in the domain (e.g.
+\statement{this animal has whiskers}, \statement{this animal has no feathers}, ...).
+The possibilities represent all cases that can be defined experimentally through the verifiable statements given by the domain.
 
-\begin{defn}
-	A \textbf{possibility} for an experimental domain $\edomain$ is a statement $x \in \tdomain$ that, when true, determines the truth value for all statements in the theoretical domain.
-Formally, $x \nequiv \contradiction$ and for each $\mathsf{s} \in \tdomain$, either $x \narrower \mathsf{s}$ or $x \ncomp \mathsf{s}$.
-The \textbf{possibilities} $X$ for $\edomain$ are the collection of all possibilities.
-\end{defn}
+There are notable properties conceptual and formal properties that can be proven from such a simple setup. The main results are the following.
 
-The possibilities are all the different cases that can be distinguished experimentally given the verifiable statements of the domain.
-If we increase or otherwise change the set of verifiable statements (e.g.
-we learn how to test the DNA of animals) then the possibilities will change as well (e.g.
-the possible animal species are refined).
-We conclude this section with the following general result.
+\textit{Each statement corresponds to a set of possibilities and logical relationships between statements correspond to set operations.} The idea that any statement in the domain (e.g.
+\statement{this animal is a mammal}) can be expressed as a disjunction of possibilities (e.g.
+\statement{this animal is a cat} $\OR$ \statement{this animal is a dog} $\OR$ ...).
+Each statement, then, can be identified by the set of possibilities that are compatible with it.
+The conjunction of two statement will be compatible with possibilities that are compatible with both statement.
+If a statement is narrower than another, then all the possibilities compatible with the first will be compatible with the second.
+We can therefore reexpress all statement relationships in term of set relationships, as in table \ref{tab:statement_set}.
 
-\begin{thrm}
-	The possibilities $X$ for an experimental domain $\edomain$ have at most the cardinality of the continuum.
-\end{thrm}
-
-The proof is simply noting that each possibility can be labeled by the truth value of the countable basis of the experimental domain.
-We cannot have more possibilities than sequences of true/false and the set of all binary sequences has the cardinality of the continuum.
-
-This means that we are never going to be able to experimentally distinguish between elements of greater cardinality.
-The set of topologically discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$, for example, has greater cardinality and therefore it will never be associated with any physically distinguishable concept.\footnote{However, a discontinuous function with up to countably many discontinuities can be seen as the limit of a sequence of continuous functions and therefore may be used as a convenient approximation of a physical concept.} All the issues with large cardinals are not something science will ever be interested in.
-It does not matter what system we are describing, what experimental techniques we are using or how clever we are.
-
-\section{Topologies and sigma-algebras}
-
-\begin{table*}
+\begin{table*}[h]
 	\centering
@@
 
-Now that we have introduced the basic mathematical structures for our general theory, we show their deep connection to other well established mathematical structures.
-We first note that each verifiable statement can be written as the disjunction of a set of possibilities.
+\textit{The experimental domain induces a natural topology on the possibilities.
+The theoretical domain induces a $\sigma$-algebra.} As the experimental domain is closed under finite conjunction and countable disjunction, the collections of sets corresponding to verifiable statements is closed under finite intersection and countable unions.
+As the domain has a countable basis, arbitrary unions become equivalent to countable ones.
+Therefore the collection of sets associated to verifiable statements is a topology on the set of possibilities.
+In the same way, the theoretical domain will correspond to a $\sigma$-algebra.
+Note that this correspond to the actual topologies commonly used.
+In the standard topology of the real numbers, for example, $(0,1)$ is an open set (a set in the topology) precisely because \statement{the value is between 0 and 1} is a verifiable statement.
+The singleton $\{1\}$ is a closed set (the complement of a set in the topology) precisely because it is not verifiable due to infinite precision, but it is falsifiable (it can be excluded experimentally).
+The verifiable part of a statement will correspond to the interior of the set.
+An undecidable statement will be one with no interior nor exterior.
+We can map topological concept readily understood in terms of experimental verifiability, as in table \ref{tab:statement_topology}.
 
-\begin{defn}
-	Let $\edomain$ be an experimental domain and $X$ its possibilities.
-We define the map $U : \edomain \rightarrow 2^X$ that for each statement $\stmt \in \edomain$ returns the set of possibilities compatible with it.
-That is, $U(\stmt)\equiv\{ x \in X \, | \, x \comp \stmt\}$.
-We call $U(\stmt)$ the \textbf{verifiable set} of possibilities associated with $\stmt$.
-\end{defn}
+\begin{table*}[h]
+	\centering
+	\begin{tabular}{p{0.075\textwidth} p{0.275\textwidth} p{0.2\textwidth} p{0.3\textwidth}}
+		& Logical concept & & Topological concept  \\ 
+		\hline 
+		$\edomain$ & (Experimental domain) & $U(\edomain)$ & (Topology) \\ 
+		$\tdomain$ & (Theoretical domain) & $U(\tdomain)$ & (Borel algebra) \\ 
+		$\stmt \in \edomain$ & (Verifiable statement) & $U(\stmt) \in U(\edomain)$ & (Open set) \\ 
+		$\tstmt \in \tdomain$ & (Theoretical statement) & $U(\tstmt) \in U(\tdomain)$ & (Borel set) \\ 		$\NOT\stmt \in \edomain$ & (Falsifiable statement) & $U(\stmt)^C  \in U(\edomain)$ & (Closed set) \\ 
+		$\ver(\stmt)$ & (Verifiable part) & $\interior(U(\stmt))$ & (Interior) \\ 
+		$\fal(\stmt)$ & (Falsifiable part) & $\exterior(U(\stmt))$ & (Exterior) \\ 
+		$\und(\stmt)$ & (Undecidable part) & $\partial U(\stmt)$ & (Boundary) 
+	\end{tabular} 
+	\caption{Correspondence between logical concepts and topological concepts.}\label{tab:statement_topology}
+\end{table*}
 
-\begin{prop}
-	A statement $\stmt \in \edomain$ is the disjunction of the possibilities in its verifiable set $U(\stmt)$. That is, $\stmt=\bigOR\limits_{x \in U(\stmt)} x$.
-\end{prop}
+\textit{The possibilities are calculated from the verifiable statements.} It is standard in physics to use the infinitely precise values as the starting point, and the finite precision interval as measurement approximations.
+In our framework it is the opposite: the finite precision verifiable statements are the starting point, and the possibilities are the limits.
+Technically, one takes the basis of the experimental domain and creates a minterm: a disjuction where each element appears once, negated or not (e.g.
+$\stmt[e]_1\OR \NOT \stmt[e]_2 \OR ...$).
+Each minterm that is not a contradiction is a possibility: something that tells us whether each element of the basis is true or note, and therefore any statement in the domain.
+This approach has various benefits.
+First, it conceptually maps better to experimental practice.
+The points of our spaces are the best description possible given a set of experimental techniques.
+If we expand the experimental techniques (e.g.
+we can measure the spin of the electron and not just its position), the possibilities are refined.
+Second, the full descriptions (i.e.
+the possibilities) and the partial descriptions (i.e.
+any other statement in the domain) are mathematically the same object.
+This makes it a lot easier to talk about partial description as they will simply be sub-domains, instead of having to define maps from sets to point and from sets to sets.
+Third, it makes it easier to understand how the assumptions that give as a particular set of points break down and what are possible alternatives.
+Assumptions will fail in narrower descriptions, not in the broader ones, so the logical relationships of the broader statements need to be preserved.
+If we start from the infinitesimal objects, we have to make modifications and then check whether the large scale behavior emerges.
+Since the framework start by defining broader statements, we are literally making small modifications on our starting points.
+That is, we keep the large scale structure fixed and see what possible small scale structure can be had.
 
-The proof is a trivial application of the disjunctive normal form of Boolean algebra.
-In fact, each possibility can be written as a minterm of a basis (i.e.
-a conjunction where each basis element appears only once either negated or not).
-Any verifiable statement can be expressed in terms of the basis, and it is a result of Boolean algebra that each logical expression can be formulated as a disjunction of minterms (i.e.
-an OR of ANDs).
-Intuitively, it is the disjunction of all cases in which the statement is true.
+\textit{The possibilities $X$ for an experimental domain $\edomain$ have at most the cardinality of the continuum.} That is, a set of objects, of cases, that can be defined experimentally cannot be bigger than the set of the real numbers.
+The proof is simply noting that each possibility can be labeled by the truth value of the countable basis of the experimental domain.
+We cannot have more possibilities than sequences of true/false and the set of all binary sequences has the cardinality of the continuum.
+This gives a clear understanding as to why objects like the set of all discontinuous functions from $\mathbb{R}$ to $\mathbb{R}$ or discussion on large cardinals are not physically relevant.
 
-Since every statement is a set of possibilities, we can re-express the statement relationships in terms of set relationships according to Table \ref{tab:statement_set}.
+\textit{The following propositions are all equivalent: all statements in an experimental domain are decidable; all theoretical statements are verifiable; all possibilities are verifiable; the possibilities form a countable basis; the topology is discrete.} This result allows to fully characterize \textbf{decidable domains}, those that contains only decidable statements.
+The discrete topology one associates with integers corresponds to the fact that we can indeed verify that \statement{there are exactly 3 apples on the table}.
+We can either verify or falsify each of the cases: all possibilities are decidable.
+This can only happen if the cases are countable, so if the set of possibilities has the cardinality of the continuum (e.g.
+the points of a manifold) we must have a qualitatively different space in terms of experimental verifiability.
 
-The closure of an experimental domain under finite conjunction and countable disjunction becomes closure under finite intersection and countable union.
-Since the basis is countable, countable union is equivalent to arbitrary union.
-In other words, the set of all verifiable sets is a topology.
+\subsection{Experimental relationships}
 
-\begin{thrm}
-	Let $X$ be the set of possibilities for an experimental domain $\edomain$.
-$X$ has a natural topology given by the collection of all verifiable sets $\mathsf{T}_X=U(\edomain)$ that is Kolmogorov and second countable.
-\end{thrm}
+Having defined experimental domain, we want to define what it means for two domains to have an inference or a causal relationship. The main result is that the two coincide.
 
-The topology is Kolmogorov (i.e.
-$T_0$) because given two possibilities, by their construction, there must be one element of the basis that is compatible with one but not the other.
-It is second countable because a basis of the experimental domain corresponds to a sub-basis of the topology.
-One can also show that the natural topology is Hausdorff if and only if all possibilities are approximately verifiable (i.e.
-each possibility is the limit of a sequence of verifiable statements).
-The hope is that we can find physically meaningful definitions for all relevant topological concepts.
+Given two domain $\edomain_Y$ and $\edomain_X$, we say there is an \textbf{inference relationship} from the first to the second if by testing a statement in the first we can infer the truth of a statement in the second.
+Formally, there is a map $\erel : \edomain_Y \to \edomain_X$ such that $\erel(\stmt_Y) \equiv \stmt_Y$.
+In that case, we say $\edomain_Y$ \textbf{depends} on $\edomain_X$, noted $\edomain_Y \subset \edomain_X$.
+If $\edomain_Y$ and $\edomain_X$ depend on each other, we say  $\edomain_Y$ is \textbf{equivalent} to $\edomain_X$, noted $\edomain_Y \equiv \edomain_X$.
 
-In the same way, the theoretical domain corresponds to a $\sigma$-algebra on the possibilities.
-
-\begin{defn}
-	Let $\tdomain$ be a theoretical domain and $X$ its possibilities.
-We define the map $A : \tdomain \rightarrow 2^X$ that for each theoretical statement $\stmt \in \tdomain$ returns the set of possibilities compatible with it.
-That is, $A(\stmt)\equiv\{ x \in X \, | \, x \comp \stmt\}$.
-We call $A(\stmt)$ the \textbf{theoretical set} of possibilities associated with $\stmt$
-\end{defn}
-
-\begin{thrm}
-	Let $X$ be the set of possibilities for a theoretical domain $\tdomain$. $X$ has a natural $\sigma$-algebra given by the collection of all theoretical sets $\Sigma_X=A(\tdomain)$.
-\end{thrm}
-
-The proof is again a simple mapping of statement operations to set operations.
-It can also be shown that the natural $\sigma$-algebra of the possibilities is the Borel algebra of their natural topology.
+Given two domain $\edomain_X$ and $\edomain_Y$, we say there is a \textbf{causal relationship} from the first to the second if by knowing which possibility $x \in X$ is true in the first we know which possibility $y \in Y$ is true in the second.
+Formally, there is a map $f : X \to Y$ such that $x \narrower f(x)$.
+Note how the direction between the domains
 
@@
 
+\bibliographystyle{plain}
 \bibliography{bibliography}

### 2020-08-04 Articles/2019-Entropy/Source/entropy-article.tex
More work on the thermo article

@@
 \begin{enumerate}
-	\item Model of a process.
-Set of all possible descriptions at all times.
-Some descriptions are narrower than others.
-Evolutions are the narrowest.
-Some descriptions are compatible.
-Each description has a unique set of compatible evolutions.
-Descriptions in narrower if the set of compatible evolution is a subset.
-	\item We can group those descriptions relative at a particular time. It is actually interval of time. States are the narrowest description at that time. How are states related through the evolution?
+	\item \emph{Experimentally testable descriptions and their logic relationship.} Here the goal is to formalize the basic structure required by a set of experimentally verifiable assertions.
+Some descriptions are more specific than others, some are compatible with each other, and so on.
+An evolution is the most specific description, one that tells us what happens at all times.
+	\item \emph{Group descriptions by time.} Here the goal is to group description at ``instants of time'', which are taken to be small intervals.
+A snapshot is the most specific description at an instant, on that tells us everything there is to know about the system at the given time.
+Determinism and reversibility are defined at this level (i.e.
+deterministic process is one where the snapshot at one time tells us the snapshots at future times).
+	\item \emph{Granularity of the description and counting evolutions.} Here the goal is to quantify how much a description is more refined than another.
+This can be done by quantifying the possible evolutions that are compatible with each (i.e.
+the more specific description will be compatible with fewer evolutions).
+The evolution entropy quantifies the difference in granularity in terms of how many questions (i.e.
+bits) are needed to go from one description to another.
+The evolution entropy cannot decrease for a deterministic process and is stationary if the process is also reversible.
 	\item Define deterministic evolution: can predict future state.
@@
 Our starting point is the \textbf{process domain} $\tdomain[P]$, which is the set that contains all possible descriptions of the system at all levels of detail at all possible times.
-Depending on the process, it will contain statements like \statement{the average volume between 1 and 2 seconds is between 3 and 4 liters} or \statement{the trajectory of the particle is $y=10 \, m - t \cdot 9.80665 \, m/s^2 $}.
+Depending on the process, it will contain statements like \statement{the average volume between 1 and 2 seconds is between 3 and 4 liters} or \statement{the trajectory of the particle is $y=10 \, m - t^2 \cdot 9.80665 \, m/s^2 $}.
 It does not matter at this point what the statements themselves are, except they must be, at least in line of principle, defined from experimentally well-defined starting points.
-Mathematically, $\tdomain[P]$ is a $\sigma$-complete Boolean algebra\footnote{A set of statements that is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).} and it is the closure of the set of experimentally verifiable statements $\edomain[P] \subseteq \tdomain[P]$.\footnote{These will form a Heyting algebra which, to be experimentally reachable, must allow a countable basis.}
+Mathematically, $\tdomain[P]$ is a $\sigma$-complete Boolean algebra\footnote{A set of statements that is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).} and it is the closure of the set of experimentally verifiable statements $\edomain[P] \subseteq \tdomain[P]$.\footnote{The set $\edomain[P]$ will form a Heyting algebra which, to be experimentally reachable, must allow a countable basis.}
 
@@
 
-\subsection{States and time}
+\subsection{Time and snapshots}
 
@@
 Note that $\tdomain_t$ is a $\sigma$-complete Boolean sub-algebra of $\tdomain[P]$ since any logical operation on statements at fixed $t$ is also another statement at fixed $t$.
-The set $X_t \subset \tdomain_t$ of the narrowest statements will give the most precise description at that time: these are the possible states of our system at that time.
+The set $X_t \subset \tdomain_t$ of the narrowest statements will give the most precise description at that time, we call these the  possible \textbf{snapshots} of our system at that time.
 
@@
 The quantities used to represent states at time $t$, then, should really be thought as averages within $\Delta t$.
-In the same way, $\tdomain_t$ will in general represent the system at the chosen level of description.
+In the same way, $\tdomain_t$ will in general represent the system at the chosen level of description, not the narrowest level possible.
 Statements about the environment or at a finer level of description (i.e.
@@
 
-We should also note that the set of states $X_t$, in general, may not be the same at all times.
-In fact $X_t$ will only include ``possible states'', those that are visited at least by one possible evolution (i.e.
-$A(x) \neq \emptyset$ for all $x \in X_t$), and it may change in time.
-For example, the set of possible states for a dissipative system will shrink as they converge to equilibiurm.
-In quantum mechanics, the possible states after a measurement are only the eigenestates and a different choice of measurement will lead to a different process with different possible evolutions.
+We should also note that the set of possible snapshots $X_t$ may not be the same at all times and may not correspond to the state of the system.
+First of all $X_t$ will only include those configurations that are compatible with least by one possible evolution (i.e.
+$A(x) \neq \emptyset$ for all $x \in X_t$), which may change in time.
+For example, the set of possible configurations for a system under dissaptive processes will shrink as they converge to equilibrium.
+In quantum mechanics, the possible configurations after a measurement are restricted to the eigenestates, and a different choice of measurement will lead to a different process with different possible evolutions.
+Secondly, the best possible description may be broader than the full state of the system (e.g.
+when external interference forces us to give only a statistical account within our $\Delta t$) or narrower (e.g.
+when there are known correlations to the other degrees of freedom of the environment or of the system itself).
 The idea of a state space of the system, independent of the process, is something we need to construct.
@@
 
-We say a process is \textbf{deterministic} over the time domains $\{X_t\}_{t \in T}$ if given the state at one time we can always predict the state at all future times.
+We say a process is \textbf{deterministic} over the time domains $\{X_t\}_{t \in T}$ if given the snapshot at one time we can always predict the snapshot at all future times.
 Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_1 \in X_{t_1}$ we can find an $x_2 \in X_{t_2}$ such that $x_1 \narrower x_2$.
@@
 Intuitively, all the evolutions that pass through $x_1$ will also pass through $x_2$.
-If $x(t)$ is the state trajectory in time, we have $A(x(t)) \subseteq A(x(t + \Delta t))$: the set of compatible trajectories must stay the same or become larger.
+If $x(t)$ is the snapshot trajectory in time, we have $A(x(t)) \subseteq A(x(t + \Delta t))$: the set of compatible evolutions must stay the same or become larger.
 
-Conversely, we say a process is \textbf{reversible} over the time domains $\{X_t\}_{t \in T}$ if given the state at one time we can always reconstruct the state at all past times.
+Conversely, we say a process is \textbf{reversible} over the time domains $\{X_t\}_{t \in T}$ if given the snapshot at one time we can always reconstruct the snapshot at all past times.
 Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_2 \in X_{t_2}$ we can find a $x_1 \in X_{t_1}$ such that $x_2 \narrower x_1$.
@@
 In this case, all the evolutions that pass through $x_2$ must have passed through $x_1$.
-If $x(t)$ is the state trajectory in time, we have $A(x(t)) \subseteq A(x(t - \Delta t))$: the set of compatible trajectories must stay the same or become smaller.
+If $x(t)$ is the snapshot trajectory in time, we have $A(x(t)) \subseteq A(x(t - \Delta t))$: the set of compatible evolutions must stay the same or become smaller.
 
 A deterministic and reversible process, then, will mean $x_1$ and $x_2$ are equivalent and that  $A(x(t)) = A(x(t + \Delta t))$: the set of evolutions remains the same.
-The fact that past and future states are equivalent $x(t) \equiv x(t + \Delta t)$ does not mean they are the same statement: it means there is an if-and-only-if relationship between them.
+The fact that past and future snapshots are equivalent $x(t) \equiv x(t + \Delta t)$ does not mean they are the same statement in terms of the description at each time: it means there is an if-and-only-if relationship between the descriptions at different times.
 For example, \statement{at time $t$ the position is $q$ and the velocity is $v$} if and only if \statement{at time $t + \Delta t$ the position is $q + v \Delta t$ and the velocity is $v$}.
@@
 
-If we limit ourselves to deterministic procseses, with these few definitions, we have already found something that ``grows bigger'' under irreversible evolution while it remains the ``the same size'' under reversible evolution.
+If we limit ourselves to deterministic processes, with these few definitions, we have already found something that ``grows bigger'' under irreversible evolution while it remains the ``the same size'' under reversible evolution.
 
@@
 
-Next step is to quantify the size of the sets of evolutions, specifically those that correspond to statements.
-Since statement at a finer level of description will put a greater constrain to the set of evolutions, quantifying the size of $A(\stmt)$ means quantifying the granularity of the description of $\stmt$.
+Next step is to quantify the size of the sets of evolutions, specifically those that correspond to each statement.
+Since statements at a finer level of description will put a greater constrain to the set of evolutions, quantifying the size of $A(\stmt)$ means quantifying the granularity of the description of $\stmt$.
 The statement \statement{at time 0 s the particle is between 0 m and 1 m} is doubly more specific than \statement{at time 0 s the particle is between 0 m and 2 m} precisely because it correspond to half the evolutions of the second statement.
 
-The naive approach would be to assign a measure to sets of $E$ and since $A(\tdomain[P])$ is a $\sigma$-algebra, it would seem we are done.
+The naive approach would be to assign a measure over $A(\tdomain[P])$: since this is is a $\sigma$-algebra over $E$, it would seem we are done.
 Unfortunately it is not that simple: a single measure can only compare objects of the same dimensionality.
 If we imagine all evolutions to be points on an $n$ dimensional manifold, if the measure gives finite values for $k$-dimensional regions it will necessarily give measure zero for all regions with lesser dimensionality and infinite measure for those with greater dimensionality.
-Coarse grained and subsystem descriptions will have different dimensionality therefore we need a way to quantify objects at all levels.\footnote{Typically one uses geometrical structure for his purpose.
+Coarse grained and subsystem descriptions will have different dimensionality as they describe different numbers of degrees of freedom, therefore we need a way to quantify objects at all levels.\footnote{Typically one uses geometrical structure for his purpose.
 In phase space, the symplectic form allows us to define areas of two-dimensional surfaces, and use those to define volumes or even-dimensional areas.
@@
 
-Therefore we will assume that we have a way to compare two statements and decide which one gives a refined description of the process.
+Therefore our starting point will be the ability to compare two statements and decide which one gives a refined description of the process.
 Formally, we have partial order $\finer$ on $\tdomain[P]$ which, given two statements, it tells us whether the description of one is \textbf{finer}, more refined, than the other.
@@
 $\stmt_1 \narrower \stmt_2$) then it is also finer than the other (i.e.
-$\stmt_1 \finer \stmt_2$) while the reverse is not necessarily true.
+$\stmt_1 \finer \stmt_2$) while the converse is not necessarily true.
 
@@
 
-We can also define the \textbf{dynamical entropy} with respect to $\stmt[u]$ the quantity $S_{\stmt[u]}(\stmt) = \log \mu_{\stmt[u]}(\stmt)$.
+We can also define the \textbf{evolution entropy} with respect to $\stmt[u]$ the quantity $S_{\stmt[u]}(\stmt) = \log \mu_{\stmt[u]}(\stmt)$.
 This can be thought as the number of bits, yes/no questions, that separate the level of description of $\stmt[u]$ from the one provided by $\stmt$.
@@
 If the process is determistic and reversible, we have $S_{\stmt[u]}(x(t)) = S_{\stmt[u]}(x(t + \Delta t))$.
-In other words, \textbf{dynamical entropy cannot decrease for a deterministic process and is stationary if the process is reversible as well}.
+In other words, \textbf{evolution entropy cannot decrease for a deterministic process and is stationary if the process is reversible as well}.
 
@@
 		\hline 
-		$S_{\stmt[u]}(\stmt)$ & Dynamical entropy & The number of bits that separate the level of description of $\stmt$ from the one provided by the unit statement $\stmt[u]$; it quantifies the number of questions needed to go from the level of description of $\stmt$ to the level of description of $\stmt[u]$  \\
+		$S_{\stmt[u]}(\stmt)$ & Evolution entropy & The number of bits that separate the level of description of $\stmt$ from the one provided by the unit statement $\stmt[u]$; it quantifies the number of questions needed to go from the level of description of $\stmt$ to the level of description of $\stmt[u]$  \\
 		\hline 
@@
 
+So far we have made no guarantees as to whether each $X_t$ contains descriptions that refer to the same physical object.
+So we need to understand what is needed to be able to define a system and its state space.
+Intuitively, the state space should provide us with all the descriptions of the system at any point in time for any process.
+We are therefore asked to create an abstraction that stitches together all possible description, all possible snapshots, about the system and only about the system.
+There are therefore a few issues we need to consider.
+
+First of all, when we say ``all process'' we can't really mean absolutely all of them, only the ones that somehow do not disrupt the system.
+For example, we can study the motion of a balloon on earth at 20 Celsius.
+This would not be possible on the surface of the sun as our balloon would not exist.
+The process has to be such that the system is recognizable to be the same, our descriptions must still make sense.
+We can imagine a space $X$ of all statement fragments like \statement{the position of the ball is between 2 and 3 meters} and a surjective map $\iota : X \times T \twoheadrightarrow X_t$ that adds \statement{at time t} for each possible time.
+The map $\iota$ must be a surjection that preserves the logical structure (i.e.
+logical operations, narrowness, compatibility, ...) so that all descriptions are only about the system and the system determines the basic logical relationships that are valid at all times (i.e.
+\statement{the position of the ball is between 2 and 3 meters} $\narrower$ \statement{the position of the ball is between 0 and 300 meters} at every $t$).
+
+Note, though, that $\iota( \cdot, t_0)$ in general is not an isomorphism (i.e.
+a bijection that preserves the logical structure) for two reasons.
+The first is that not all states will be available at every moment in every process.
+As we said before, if the process is dissipative the set of possible states will become smaller.
+In that case, $\iota$ will map all the inaccessible states to $\contradiction$, as they will be impossible.
+Therefore $\iota$ is not in general a bijection.
+The second reason is that the statements in $X$ represent only descriptions about the system itself and nothing else.
+In a particular process, however, knowing the state of one system might tell us something about other system as well.
+The presence of correlations/coupling may mean that the precise knowledge of the degrees of freedom of the system may allow us to tell something about the environment or the internal structure (i.e.
+microstate).
+Mathematically, $\iota$ may map to statements that are narrower than the original (e.g.
+we can have $\iota($\statement{position of A is 1 meter}$, t_0) \equiv$ \statement{at time $t_0$ the position of A is 1 meter and the position of B is -1 meter}).
+
+States coincide with snapshots of the system only in the case that the system is uncorrelated with environment and it does constrain the internal dynamics.
+This is also the case when the statements are broadest, which means it the case when the process entropy is maximized.
+We therefore define the \textbf{state entropy} as the maximum of all possible evolution entropies one can find for the given description of the system in all possible processes at all times.
+
+In this context, entropy maximization is a necessary step to be able to define a system.
+If we are not able to find a process that, at least in some cases with good approximation, makes the evolution of the system independent from the environment and from the internal dynamics, we cannot talk about a system to begin with.
+In other words, if we are not able to isolate the system so that it can be studied independently, we are not going to be able to assign it a state, we are not able to construct a state space.
+
+It is then useful to start studying pro
+
+The statements present in our state space $X$, then, represent theIf $\iota$ can map to statements of different granularity, we cannot assign a unique evolution entropy to the statements of $X$. Yet,
+
+ In our framework, we need some type of homomorphism between $\tdomain_t$ and $\tdomain_{t + \Delta t}$ that preserves narrowness and fineness, that tells us the system is the same.
+
 So far we have only characterized how descriptions at different times can be related to each other.
 We now want to group descriptions that belong together, that describe a single system as it is subjected to different processes.
+In other words: we want to start talking about \textit{states of the system}, not just descriptions or snapshots.
 In the simplest form, we prepare the system in an initial state, let it evolve and look at the final state.
@@
 Yet, it also tells us there are a few things that we have to worry about.
-
-First of all, the process can't really be any process.
-We can study the motion of a balloon on earth at 20 Celsius.
-This would not be possible on the surface of the sun as our balloon would not exist.
-The process has to be such that the system is recognizable to be the same.
-In our framework, we need some type of homomorphism between $\tdomain_t$ and $\tdomain_{t + \Delta t}$ that preserves narrowness and fineness, that tells us the system is the same.
 

### 2020-08-04 Articles/2020-Refactoring/Source/refactoring-article.tex
More work on the refactoring article

@@
 
-Axioms, definitions and justifications are included in the mathematical section, with the rest of the formal elements. They should be seen as integral part of the framework.
+Axioms, definitions and justifications are included in the mathematical section, with the rest of the formal elements.
+The informal elements should be seen as integral part of the framework as we should never be allowed to ``assume'' something that was not fully backed by a scientific explanation.
 
-\section{Basic formal structures?}
+\section{An overview of the mathematical structures}
 
 The starting point of our framework is the \textbf{principle of scientific objectivity} which states \statement{Science is universal, non-contradictory and evidence based}.
-Every physical theory, then, must give us a set of logically sound statements linked with some notion of experimental verification.
-The content of the statements, what can be verified and how, will vary from theory to theory, but logical consistency and verifiability must be common to all of them.
+Every physical theory, then, must give us a set of logically sound statements together with some notion of experimental verification.
+While the content of the statements, what can be verified and how, will vary from theory to theory, the rules dictated by logical consistency and experimental verifiability must be always followed.
 Therefore we will have two set of axioms: the \textbf{logical consistency axioms} and the \textbf{experimental verifiability axioms}.
@@
 Each context has to be such that the meaning of each statement has no ambiguity and therefore it is associated with a single, well-defined truth value.
-This is, in general, not known: it cannot be proven or deduced.
+This value is, in general, not known, it cannot be proven or deduced.
 Ultimately, it is the role of experimentation to find it.
@@
 We can also also define \textbf{decidable statements} as the ones that can be both verified and falsified experimentally (i.e.
-their negation is a verifiable statement).
+their negation is also a verifiable statement).
 One can prove that the finite disjunction and the finite conjunction of decidable statements is also decidable.
@@
 
-\subsection{Domains and possibilities?}
+\subsection{Experimental domains and possibilities}
 
@@
 Even when given an arbitrarily long time, a countable set of statements is the most that can be verified.
-However, due to the logical relationship between statements, one does not need to test all of them: if \statement{the mass of the photon is less than $10^{-6}$ eV} is experimentally verified, there is no need to test whether \statement{the mass of the photon is greater than $10^{-6}$ eV}.
+However, one does not need to test all of statements within a set as some results can be inferred: if \statement{the mass of the photon is less than $10^{-6}$ eV} is experimentally verified, there is no need to test whether \statement{the mass of the photon is greater than $10^{-6}$ eV}.
 
@@
 Instead, if $\stmt=$\statement{the mass of the photon is less than $10^{-17}$ eV}, we still have $\ver(\stmt) \equiv \stmt$ because it is a verifiable statement, but we will have $\fal(\stmt) \equiv$\statement{the mass of the photon is more than $10^{-17}$ eV} while $\und(\stmt) \equiv$\statement{the mass of the photon is exactly $10^{-17}$ eV}.
-The issue is that, due to finite precision, if the mass of the photon where exactly $10^{-17}$ eV, our confidence interval would only include the region below that number, so we would never be able to exclude it.
+The issue is that, due to finite precision, if the mass of the photon where exactly $10^{-17}$ eV, our confidence interval would always include the region below that number, so we would never be able to exclude it.
 So, the test would not terminate in that particular case.
@@
 
-There are notable properties conceptual and formal properties that can be proven from such a simple setup. The main results are the following.
+There are notable conceptual and formal properties that can be proven from such a simple setup. The main results are the following.
 
-\textit{Each statement corresponds to a set of possibilities and logical relationships between statements correspond to set operations.} The idea that any statement in the domain (e.g.
+\textit{Each statement in a domain corresponds to a unique set of possibilities and logical relationships between statements correspond to set operations.} The idea that any statement in the domain (e.g.
 \statement{this animal is a mammal}) can be expressed as a disjunction of possibilities (e.g.
@@
 Each statement, then, can be identified by the set of possibilities that are compatible with it.
-The conjunction of two statement will be compatible with possibilities that are compatible with both statement.
-If a statement is narrower than another, then all the possibilities compatible with the first will be compatible with the second.
-We can therefore reexpress all statement relationships in term of set relationships, as in table \ref{tab:statement_set}.
+The conjunction of two statements will be compatible with possibilities that are compatible with both statements, therefore conjunction corresponds to intersection.
+If a statement is narrower than another, then all the possibilities compatible with the first will be compatible with the second, therefore the narrowness relationship corresponds to subset relationship.
+We can re-express all statement relationships in term of set relationships, as in table \ref{tab:statement_set}.
 
@@
 In the same way, the theoretical domain will correspond to a $\sigma$-algebra.
-Note that this correspond to the actual topologies commonly used.
+Note that these correspond to the actual topologies commonly used.
 In the standard topology of the real numbers, for example, $(0,1)$ is an open set (a set in the topology) precisely because \statement{the value is between 0 and 1} is a verifiable statement.
@@
 An undecidable statement will be one with no interior nor exterior.
-We can map topological concept readily understood in terms of experimental verifiability, as in table \ref{tab:statement_topology}.
+Topological concepts can readily be understood in terms of experimental verifiability, as in table \ref{tab:statement_topology}.
 
@@
 $\stmt[e]_1\OR \NOT \stmt[e]_2 \OR ...$).
-Each minterm that is not a contradiction is a possibility: something that tells us whether each element of the basis is true or note, and therefore any statement in the domain.
+Each minterm that is not a contradiction is a possibility: something that tells us whether each element of the basis is true or not, and therefore any statement in the domain.
 This approach has various benefits.
@@
 This makes it a lot easier to talk about partial description as they will simply be sub-domains, instead of having to define maps from sets to point and from sets to sets.
-Third, it makes it easier to understand how the assumptions that give as a particular set of points break down and what are possible alternatives.
+Third, it makes it easier to understand how the assumptions that give us a particular set of points break down and what are possible alternatives.
 Assumptions will fail in narrower descriptions, not in the broader ones, so the logical relationships of the broader statements need to be preserved.
@@
 \textit{The following propositions are all equivalent: all statements in an experimental domain are decidable; all theoretical statements are verifiable; all possibilities are verifiable; the possibilities form a countable basis; the topology is discrete.} This result allows to fully characterize \textbf{decidable domains}, those that contains only decidable statements.
-The discrete topology one associates with integers corresponds to the fact that we can indeed verify that \statement{there are exactly 3 apples on the table}.
+The discrete topology one associates with integers corresponds to the fact that statements like \statement{there are exactly 3 apples on the table} are decidable.
 We can either verify or falsify each of the cases: all possibilities are decidable.
@@
 
-Having defined experimental domain, we want to define what it means for two domains to have an inference or a causal relationship. The main result is that the two coincide.
+Having defined experimental domain, we proceed to define what it means for two domains to be experimentally related.
+There are two types of relationships we will consider: inference and causal.
+The main result is that the two coincide.
 
@@
 Formally, there is a map $\erel : \edomain_Y \to \edomain_X$ such that $\erel(\stmt_Y) \equiv \stmt_Y$.
-In that case, we say $\edomain_Y$ \textbf{depends} on $\edomain_X$, noted $\edomain_Y \subset \edomain_X$.
+In that case, we say $\edomain_Y$ \textbf{depends} on $\edomain_X$, noted $\edomain_Y \subseteq \edomain_X$.
 If $\edomain_Y$ and $\edomain_X$ depend on each other, we say  $\edomain_Y$ is \textbf{equivalent} to $\edomain_X$, noted $\edomain_Y \equiv \edomain_X$.
@@
 Formally, there is a map $f : X \to Y$ such that $x \narrower f(x)$.
-Note how the direction between the domains
+Note how the direction between the domains is different between causal and inference relationship: the inference proceeds from effect to cause.
+Given these definition, one can prove the following.
+
+\textit{Causal and inference relationships are equivalent.} That is, if there exists a causal relationship between $\edomain_X$ and $\edomain_Y$ then there is an inference relationship between $\edomain_Y$ and $\edomain_X$ and vice-versa.
+For example, if the temperature of the water determines its density, then by measuring the density we can infer the temperature of the water.
+We say, then, that there is an \textbf{experimental relationship} between the two domains.
+Note the relationship is directional: the water density does not determine the temperature because water is most dense at 4 Celsius.
+In the same way, a measurement on temperature is not in general equivalent to a measurement of density because it can tell us whether we are above or below 4 Celsius, which can't be done with a measurement on density.
+These are the types of subtleties that a mathematical framework that closely maps to physical concepts helps brings to light.
+
+\textit{Causal relationship are topologically continuous functions.} Given a causal relationship $f : X \to Y$, for every verifiable statement in $\edomain_Y$ we can infer an equivalent statement in $\edomain_X$, which must be verifiable since it is equivalent to a verifiable statement.
+In terms of the topology, $f^{-1}(U)$ must be an open set if $U$ is open, therefore $f$ must be continuous.
+This tells us why in physics functions are ``well-behaved''.
+It is not a matter of convenience: they must be so or they would break experimental verifiability.
+
+\textit{Domain equivalence is isomorphism in the category.} If we were to add additional mathematical structure to our domain, given that the domain equivalence preserves statement equivalence, the additional mathematical structure will be preserves as well.
+Therefore, if we had a manifold and a differentiable structure, domain equivalence would be a diffeomorphism; if we had a vector space structure, domain equivalence would be a linear transformation and so on.
+This again tells us when and why these mathematical concepts are important in physics.
+
+\subsection{Combined domains}
+
+Another general operation we can define is domain combination, which takes two or more domains and merges them into one that contains all of them.
+
+Given two domains $\edomain_X$ and $\edomain_Y$ we define the \textbf{combined domain} $\edomain_X \times \edomain_Y$ as the one that contains all the verifiable statements that can be constructed from both $\edomain_X$ and $\edomain_Y$.
+For example, if one domain describes the shape of an object and the other its color, the combined domain will contain descriptions of both.
+Formally, $\edomain_X \times \edomain_Y$ is the closure under finite conjunction and countable disjunction.
+This definition can be extended to the case of countably many domains.
+
+The goal is to understand how the possibilities depend on the type of relationship we have between the domains.
+We therefore introduce the following definitions.
+Two domains are said to be \textbf{independent} if taking a statement from each domain always gives a set of independent statements.
+Formally, $\stmt_X \indep \stmt_Y$ for all $\stmt_X \in \edomain_X$ and $\stmt_Y \in \edomain_Y$.
+Two domains are said to be \textbf{incompatible} if all statements from one are incompatible with all statements from the other, with the exception of tautologies.
+Formally, $\stmt_X \ncomp \stmt_Y$ for all $\stmt_X \in \edomain_X \setminus \{\tautology\}$ and $\stmt_Y \in \edomain_Y \setminus \{\tautology\}$.
+Now a list of consequences for these definitions.
+
+\textit{The set of possibilities of the combined domain is a subset of the scalar product of the possibilities of the individual domains.} Intuitively, a complete description of the individual domains is equivalent to the complete description of both domain combined.
+Not all combinations, though, may be possible because there may be a dependence between the two domains.\footnote{It is an open question what are the possible subsets of the scalar product are allowed for a combined domain.}
+
+\textit{If $\edomain_Y$ depends on $\edomain_X$, then the combined domained is $\edomain_X$.} Intuitively, if $\edomain_Y$ depends on $\edomain_X$ (e.g.
+the domain for water density depends on the one for water temperature), then $\edomain_X$ already contains all statements in $\edomain_Y$ and all its possibilities, so nothing is added by the combination.
+
+\textit{If $\edomain_X$ is independent of $\edomain_Y$, then the combined possibilities coincides with the scalar product.} If the domains are independent (e.g.
+vertical position and horizontal position), so will be their possibilities.
+Each possibility from the first can be combined with any possibility of the second, so all combinations are allowed.
+The natural topology will correspond to the product topology even in the case of countably many domains.\footnote{Note that it will naturally correspond to the product topology, and not the box topology, because of closure under finite conjunction.}
+
+\textit{If $\edomain_X$ is incompatible to $\edomain_Y$, then the combined possibilities coincides with the disjoint union of the possibilities.} If the domains are incompatible (e.g.
+descriptions for a plant and descriptions for an animal), the possibilities will also describe incompatible cases.
+So, conceptually, the each combined possibility will be a possibility from one of the two domains.
+The details are a bit more involved and they are available in the appendix.
 

### 2020-08-21 Articles/2019-Entropy/Source/entropy-article.tex
Up to entropy maximization

@@
 
-The overall goal is to describe a generic system as it evolves in time by defining a minimal set of concepts that such a setup must define.
-We want to find notions that can be applied, for example, to thermodyamic and dynamical systems alike.
+The overall goal is to describe a generic system as it evolves in time by defining a minimal set of concepts that such a setup must define, and understand the tacit assumptions underlying state spaces and evolution laws.
+We want to find notions that can be applied in different settings, being it thermodyamics and/or dynamical systems alike.
 Then we want to derive general results based on such definitions.
@@
 It does not matter at this point what the statements themselves are, except they must be, at least in line of principle, defined from experimentally well-defined starting points.
-Mathematically, $\tdomain[P]$ is a $\sigma$-complete Boolean algebra\footnote{A set of statements that is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).} and it is the closure of the set of experimentally verifiable statements $\edomain[P] \subseteq \tdomain[P]$.\footnote{The set $\edomain[P]$ will form a Heyting algebra which, to be experimentally reachable, must allow a countable basis.}
+Mathematically, $\tdomain[P]$ is a $\sigma$-complete Boolean algebra\footnote{A set of statements that is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).} and it is the closure of the set of experimentally verifiable statements $\edomain[P] \subseteq \tdomain[P]$.\footnote{The set $\edomain[P]$ will form a Heyting algebra which, to be experimentally reachable, must allow a countable basis.
+The closure $\tdomain[P]$ is in terms of negation and countable disjunction.}
 
@@
 As with narrowness, compatibility can be in terms of different ranges of the same quantities, of different quantities and of constraints at equal or different times.
+
+Finally, we want to capture whether the truth of one or more statements incluences the truth of others.
+For example, we want to say that \statement{the temperature of the water in the glass is 3.98 C} is \textbf{independent} from \statement{the volume of the water in the glass is 10 $cL$} but is not independent from \statement{the density of the water in the glass is 1 $g/cm^3$}.
+We note $\stmt_1 \indep \stmt_2$ when two statements are independent.
 
@@
 	\hline 
+	$\indep$ & independent & Two statements are independent  $\stmt_1 \comp \stmt_2$ if the truth of one does not influence the truth of the other \\ 
+\hline 
 \end{tabular} 
@@
 As the process extends over time, we want to organize the statements in our process domain into partial domains, one for each moment in time.
-Given a \textbf{time parameter} $t \in T \subseteq \mathbb{R}$, we can imagine to carve a subdomain $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
+Given a \textbf{time parameter} $t \in T \subseteq \mathbb{R}$, we can imagine to carve a \textbf{system domain} $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
+In particular, we define a \textbf{one-step processes} one that only has an \textbf{initial domain} $\tdomain_{t_0}$ and a \textbf{final domain} $\tdomain_{t_1}$.
 Note that $\tdomain_t$ is a $\sigma$-complete Boolean sub-algebra of $\tdomain[P]$ since any logical operation on statements at fixed $t$ is also another statement at fixed $t$.
@@
 We should also note that the set of possible snapshots $X_t$ may not be the same at all times and may not correspond to the state of the system.
-First of all $X_t$ will only include those configurations that are compatible with least by one possible evolution (i.e.
+First of all $X_t$ will only include those configurations that are compatible with at least by one possible evolution (i.e.
 $A(x) \neq \emptyset$ for all $x \in X_t$), which may change in time.
@@
 when there are known correlations to the other degrees of freedom of the environment or of the system itself).
-The idea of a state space of the system, independent of the process, is something we need to construct.
-Finding the requirements needed for that construction, the hidden assumptions, is precisely the issue we need to address.
+In other words, the state space of the system (i.e.
+the set of all possible complete descriptions of the system and only about that system, independently of the process at hand) is something we'll construct later.
 
-We say a process is \textbf{deterministic} over the time domains $\{X_t\}_{t \in T}$ if given the snapshot at one time we can always predict the snapshot at all future times.
+We say a process is \textbf{deterministic} over the system domains $\{X_t\}_{t \in T}$ if given the snapshot at one time we can always predict the snapshot at all future times.
 Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_1 \in X_{t_1}$ we can find an $x_2 \in X_{t_2}$ such that $x_1 \narrower x_2$.
@@
 
-Conversely, we say a process is \textbf{reversible} over the time domains $\{X_t\}_{t \in T}$ if given the snapshot at one time we can always reconstruct the snapshot at all past times.
+Conversely, we say a process is \textbf{reversible} over the system domains $\{X_t\}_{t \in T}$ if given the snapshot at one time we can always reconstruct the snapshot at all past times.
 Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_2 \in X_{t_2}$ we can find a $x_1 \in X_{t_1}$ such that $x_2 \narrower x_1$.
@@
 
-If we limit ourselves to deterministic processes, with these few definitions, we have already found something that ``grows bigger'' under irreversible evolution while it remains the ``the same size'' under reversible evolution.
+If we limit ourselves to deterministic processes, with these few definitions, we have already found that the set of possible evolutions ``grows bigger'' under irreversible evolution while it remains the ``the same size'' under reversible evolution.
+This is the core idea of the second law.
 
@@
 		\hline 
-		$\tdomain_t \subset \tdomain[P]$ & Time domain & The set of descriptions at a given time \\ 
+		$\tdomain_t \subset \tdomain[P]$ & System domain & The set of descriptions at a given time \\ 
 		\hline 
-		$X_t \subset \tdomain_t$ & Possible states & The set of descriptions that give a full account at time $t$; they correspond to the narrowest statements within $\tdomain_t$ \\ 
+		$X_t \subset \tdomain_t$ & Possible snapshots & The set of descriptions that give a full account at time $t$; they correspond to the narrowest statements within $\tdomain_t$ \\ 
 		\hline 
-		$x(t) \narrower x(t + \Delta t)$ & Determinism & Given the state at one time we can predict the state at future times \\ 
+		$x(t) \narrower x(t + \Delta t)$ & Determinism & Given the snapshot at one time we can predict the snapshot at future times \\ 
 		\hline 
-		$x(t) \narrower x(t - \Delta t)$ & Reversibility & Given the state at one time we can reconstruct the state at past times \\ 
+		$x(t) \narrower x(t - \Delta t)$ & Reversibility & Given the snapshot at one time we can reconstruct the snapshot at past times \\ 
 		\hline 
 	\end{tabular} 
-	\caption{Time domain, states and determinism}
+	\caption{System domain, snapshots, determinism and reversibility}
 	\label{table:states}
@@
 If we consider all the evolutions compatible with \statement{the coin is fair}, we must have that half of them are compatible with \statement{the coin is head} while the other are compatible with \statement{the coin is tails}.
-Therefore we have $P(\stmt_2 | \stmt_1) = \mu_{\stmt_1}(\stmt_1 \AND \stmt_2)$ where $\stmt_1 \AND \stmt_2$ is the logical conjunction (AND) between the two statements.
+Therefore we have that the \textbf{probability} $P(\stmt_2 | \stmt_1) = \mu_{\stmt_1}(\stmt_1 \AND \stmt_2)$ where $\stmt_1 \AND \stmt_2$ is the logical conjunction (AND) between the two statements.
 In other words, the probability quantifies the ratio of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$.
+
+We can study a one-step process by tracking the probability $P(\stmt_1 | \stmt_0)$ that the final description $\stmt_1$ is true given that the initial description $\stmt_0$ is true, which is simply the fractions of evolutions starting in $\stmt_0$ (i.e.
+are compatible to $\stmt_0$)) that end in $\stmt_1$.
+The fact that \textbf{probability theory can be used to fully characterize all one-step processes}, the simplest processes, tells us why it is so important in physics.
 
@@
 		\hline 
-		$S_{\stmt[u]}(\stmt)$ & Evolution entropy & The number of bits that separate the level of description of $\stmt$ from the one provided by the unit statement $\stmt[u]$; it quantifies the number of questions needed to go from the level of description of $\stmt$ to the level of description of $\stmt[u]$  \\
+		$S_{\stmt[u]}(\stmt)$ & Evolution entropy & Defined as $\log \mu_{\stmt[u]}(\stmt)$ represents the number of bits that separate the level of description of $\stmt$ from the one provided by the unit statement $\stmt[u]$; it quantifies the number of questions needed to go from the level of description of $\stmt$ to the level of description of $\stmt[u]$  \\
 		\hline 
-		$P(\stmt_2 | \stmt_1)$ & Probability & Quantifies quantifies the ratio of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$  \\ 
+		$P(\stmt_2 | \stmt_1)$ & Probability & Defined as $\mu_{\stmt_1}(\stmt_1 \AND \stmt_2)$ quantifies the ratio of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$  \\ 
 		\hline 
@@
 
-\subsection{Systems, equilibria and state spaces}
+\subsection{States and system independence}
 
-So far we have made no guarantees as to whether each $X_t$ contains descriptions that refer to the same physical object.
+So far we have made no guarantees as to whether each $\tdomain_t$ contains descriptions that refer only to a particular physical object.
 So we need to understand what is needed to be able to define a system and its state space.
-Intuitively, the state space should provide us with all the descriptions of the system at any point in time for any process.
-We are therefore asked to create an abstraction that stitches together all possible description, all possible snapshots, about the system and only about the system.
-There are therefore a few issues we need to consider.
+Intuitively, the state space should provide us with all the descriptions of the system, and only of the system, at any point in time for any process.
+We are therefore asked to create an abstraction that stitches together all possible descriptions, all possible snapshots, about the system and only about the system, which means there are a few issues we need to consider.
 
-First of all, when we say ``all process'' we can't really mean absolutely all of them, only the ones that somehow do not disrupt the system.
+First of all, when we say ``all processes'' we can't really mean absolutely all of them, but only the one for which the system is well defined.
 For example, we can study the motion of a balloon on earth at 20 Celsius.
@@
 The process has to be such that the system is recognizable to be the same, our descriptions must still make sense.
-We can imagine a space $X$ of all statement fragments like \statement{the position of the ball is between 2 and 3 meters} and a surjective map $\iota : X \times T \twoheadrightarrow X_t$ that adds \statement{at time t} for each possible time.
+We can imagine a \textbf{state domain} $\tdomain$ of all statement fragments like \statement{the position of the ball is between 2 and 3 meters} together with a surjective map $\iota : \tdomain \times T \twoheadrightarrow \tdomain_t$ that adds \statement{at time t} for each possible time.
 The map $\iota$ must be a surjection that preserves the logical structure (i.e.
-logical operations, narrowness, compatibility, ...) so that all descriptions are only about the system and the system determines the basic logical relationships that are valid at all times (i.e.
-\statement{the position of the ball is between 2 and 3 meters} $\narrower$ \statement{the position of the ball is between 0 and 300 meters} at every $t$).
+logical operations, verifiability, narrowness, compatibility, ...) so that the basic logical relationships determined by the system itself are are valid at all times (i.e.
+\statement{at time t the position of the ball is between 2 and 3 meters} $\narrower$ \statement{at time t the position of the ball is between 0 and 300 meters} for every $t$).\footnote{We stress that the mathematical requirements are necessary but not sufficient.
+While the same system must map to the same logical structure but the same logical structure does not necessarily map to the same system (e.g.
+two distinct system can be represented by the same mathematical structure).}
 
@@
 Therefore $\iota$ is not in general a bijection.
-The second reason is that the statements in $X$ represent only descriptions about the system itself and nothing else.
+The second reason is that the statement fragments in $\tdomain$ represent only descriptions about the system itself and nothing else.
 In a particular process, however, knowing the state of one system might tell us something about other system as well.
@@
 microstate).
-Mathematically, $\iota$ may map to statements that are narrower than the original (e.g.
+Mathematically, $\iota$ may map to statements that are narrower than the original fragment (e.g.
 we can have $\iota($\statement{position of A is 1 meter}$, t_0) \equiv$ \statement{at time $t_0$ the position of A is 1 meter and the position of B is -1 meter}).
 
-States coincide with snapshots of the system only in the case that the system is uncorrelated with environment and it does constrain the internal dynamics.
-This is also the case when the statements are broadest, which means it the case when the process entropy is maximized.
-We therefore define the \textbf{state entropy} as the maximum of all possible evolution entropies one can find for the given description of the system in all possible processes at all times.
+In physics terms, the construction of the state domain requires us to be able to identify and separate the system from all other systems and the environment.
+We can talk about a book and its state because we can, in many situations, describe and manipulate the book without describing or manipulating other books, the bookcase or the table.
+If $\tdomain_{t}$ is the domain of our system at time $t$ and $\tdomain[O]_t$ is the domain for another system, we say that the two domains are \textbf{independent} if, taken any two statements $\stmt_d \in \tdomain_{t}$ and $\stmt_o \in \tdomain[O]_{t}$ are independent.
+In this case there are no correlations: every possible pair of snapshots of each domain is a possible snapshot for the combined system.
+Therefore if we take $\stmt[u]_d \in \tdomain_{t}$ and $\stmt[u]_o \in \tdomain[O]_{t}$ to be unit statements for the respective measure, we have:
+\begin{equation}\label{entropy_sums}
+\begin{aligned}
+	\mu_{\stmt[u]_d \AND \stmt[u]_o}(\stmt_d \AND \stmt_o)=\mu_{\stmt[u]_d}(\stmt_d) \mu_{\stmt[u]_o}(\stmt_o) \\
+S_{\stmt[u]_d \AND \stmt[u]_o}(\stmt_d \AND \stmt_o)=S_{\stmt[u]_d}(\stmt_d) + S_{\stmt[u]_o}(\stmt_o)
+\end{aligned}
+\end{equation} 
+That is, \textbf{for independent systems the process entropy can be seen as the sum of the entropy of each system}.\footnote{Note the direction in which the construction unfolds.
+We are not defining the entropy on the single systems and then, if they are isolated, the entropy sums.
+It's the opposite: the entropy is initially defined on the joint system for a particular process as the evolutions describe all systems.
+If the system is independent of all others, the evolutions are grouped such that their measure factorizes, and we can assign a unique measure to each statement of the independent system, as this will not depend on the description of other systems.
+If we cannot find this situation, the entropy of the system cannot be defined uniquely on the separate system, but only on the joint.
+In physics terms, we are unable to treat the description as set of independent degrees of freedom (e.g.
+describing only the pressure of a gas without the volume, or only the position without the momentum).}
 
-In this context, entropy maximization is a necessary step to be able to define a system.
-If we are not able to find a process that, at least in some cases with good approximation, makes the evolution of the system independent from the environment and from the internal dynamics, we cannot talk about a system to begin with.
-In other words, if we are not able to isolate the system so that it can be studied independently, we are not going to be able to assign it a state, we are not able to construct a state space.
+We say we have an \textbf{independent system} if the system is independent from all other systems.
+In this case, the snapshot is a description of the system and only of the system.
+The state domain, then, represent descriptions of the system when it is independent.
+A \textbf{state} $x$ of a system is a snapshot in these conditions and the \textbf{state space} is the set of all possible states.\footnote{There is still the added complication that, even if the system is independent, not all states may be accessible.
+For example, the thermodynamic states may change depending on the constraint applied to the system (e.g.
+presence or absence of a magnetic field); in quantum mechanics, the accessible states after a spin measurement will only be the ones along the measured direction.
+So, in general, the state space stitches together all the possible independent snapshots at all times over all possible processes.} In the same way, the process entropy becomes a property of the system itself and we can define the \textbf{state entropy} $S_{\stmt[u]} : X \to \mathbb{R}$ as the process entropy in these conditions, where $\stmt[u] \in X$ is a state taken as reference.
+That is:
+\begin{equation}
+	S_{\stmt[u]}(x) = S_{\iota(\stmt[u], t_0)}(\iota( x, t_0))_{indep}
+\end{equation}
 
-It is then useful to start studying pro
+We say $\xi^a : X \to \mathbb{R}$ are a set of state variables if they uniquely identify a state.
+That is, we can find an isomorphism (a bijection that preserve the logical structure) $\phi : \mathbb{R}^n \to X$ such that $\phi(\xi^a(x)) = x$ for all $x \in X$.
+Then each state space comes equipped with a \textbf{state function} $S=S(\xi^a)$.
+We'll see that, for thermodynamic systems, this will correspond to the usual thermodynamic entropy.
+We define a \textbf{purely mechanical} system one for which the state entropy is the same across all states (i.e.
+$S=k$ where $k$ is a constant, typically chosen to be zero).
+We'll see that classical phase space and quantum state spaces are state spaces of purely mechanical systems.
 
-The statements present in our state space $X$, then, represent theIf $\iota$ can map to statements of different granularity, we cannot assign a unique evolution entropy to the statements of $X$. Yet,
+Since we can only properly talk about a system when it is independent from the other, what can we do in the other cases?
+The idea that we can still talk about a system in that case if we are able, at least in line of principle, to decouple the system from all others while not changing the nature of the system itself.
+For example, we can talk about a molecule of a gas because we have a way to take such molecule out of the gas.
+The construction of a state space of a system, then, also requires telling us what all the possible correlated description will go when the system is made independent.
+Mathematically, we need a deterministic one-step process that takes any initial snapshot of the system to a final state, and does so without disrupting the system (i.e.
+the relaxation process must take each state to itself).
+We call this process \textbf{relaxation}.\footnote{The state space is invariant over the relaxation.
+In other word, the state space is a symmetry of the relaxation process.}
 
- In our framework, we need some type of homomorphism between $\tdomain_t$ and $\tdomain_{t + \Delta t}$ that preserves narrowness and fineness, that tells us the system is the same.
+If $x_0$ is the initial snapshot and $x_1$ is the final state, given that the relaxation process is deterministic, we have $x_1= x_1(x_0)$ and 
+\begin{equation}
+S_{\iota(\stmt[u], t_1)}(x_0) \leq S_{\iota(\stmt[u], t_1)}(x_1) = S_{\iota(\stmt[u], t_1)}(\iota( x, t_1)) = S_{\stmt[u]}(x)
+\end{equation}
+The process entropy increases and is maximized when the system is independent.\footnote{Mathematically, the relaxation process in needed to compare the process entropies for the correlated descriptions.
+As the same description can be more or less correlated in different cases, the independent ones work as a reference and we need a way to relate to them.}
 
-So far we have only characterized how descriptions at different times can be related to each other.
-We now want to group descriptions that belong together, that describe a single system as it is subjected to different processes.
-In other words: we want to start talking about \textit{states of the system}, not just descriptions or snapshots.
-In the simplest form, we prepare the system in an initial state, let it evolve and look at the final state.
-We want to characterize the state space of all possible initial and final states under all process and characterize the process.
-Our framework already tells us that if the process is deterministic, the process can be characterized as a map $x(t+\Delta t) = f(x(t))$.
-If the process is non-deterministic, the evolutions for each state may split, and the probability $P(x(t + \Delta t) | x(t))$ can be used to characterize the fraction of evolutions that start in $x(t)$ and end in $x(t+\Delta t)$.
-Yet, it also tells us there are a few things that we have to worry about.
+Note that all snapshots that relax to the same state will have a process entropy lower or equal to the state entropy.
+Therefore \textbf{the state entropy is the largest possible process entropy} for which the description given by the state applies.
+This does not simply represent a subjective information theoretic consideration (e.g.
+we know only about the system); it is a physical consideration (e.g.
+knowledge about one system tells us nothing about the others).
+As we make the system more independent, there are more possible evolutions the combined system can have.
+Therefore in this framework \textbf{entropy maximization is equivalent to requiring system independence}.
+Entropy maximization, more than a law of physics, is a requirement to properly define the state space of a system, a requirement for doing physics.
+As such, we are therefore biased towards looking for such situations.
 
-Second we want to make sure that the state space of the system really tells us only about the system itself.
-Within a specific process, for example, saying \statement{at time $t$, the velocity of object $A$ is $v$} may also mean that \statement{at time $t$, the velocity of object $B$ is $v$} because they are tied together.
-We say two systems are \textbf{independent} at a particular time if the description of one system does not constrain the description of the other.
-In terms of measure $\mu_{\stmt[u]_a \AND \stmt[u]_b}(x_a \AND x_b) = \mu_{\stmt[u]_a}(x_a) \mu_{\stmt[u]_b}(x_b)$ must factorize.
-If we want our state space to describe only the system at hand, then we are assuming that states are defined on the assumption that the system is independent at least from all other systems with the chosen processes.
+\subsection{Thermodynamic processes}
 
+% DO WE NEED INTERNAL STATE?
 On the other side, a state of a system is never really the description of the whole system, but of the part that is relevant to the process at hand.
@@
 We say a system is in \textbf{equilibrium} if the state tells us nothing about the
-
-We stress that the independence cannot be, in general, independent throughout the whole process or there cannot be interactions between the systems.
-By the same token, the finer dynamics may play a part throughout the process.
-The constraint is only on the initial and final states, where the state space is being constructed.
-We also stress that these constraints make the state definition contingent upon the type of processes one studies.
-If we want to study a process where two systems are not independent, one that depends on the finer dynamics, our state definition will have to change to accommodate.
-
- In all processes we are considering, then, the system must be indAny other system would be independent If we found other systems that were always correlated For example, if  There can't be other systems that participate in the process whose state
-
-We should caution that
-
-The simplest setting is the follow
-
- into Our goal now is to construct state spaces for a system.
-
- Consequently any statement that corresponds to finitely many evolutions (e.g.
-\statement{the full trajectories is either $y=t \cdot 1 m/s$ or $y=-t \cdot 1 m/s$}) will be comparable to all other statements.
-
-Also note that in our ideal gas example one of the quantities was determined with infinite precision.
-If we gave a range for all quantities, we would compare areas in that space, which we can always do.
-Statements with finite precision measurements are the only ones we can experimentally verify.
-So, we will require that all experimentally verifiable statements are comparable.
-We call \textbf{experimental domain} for the process the subset $\edomain[P] \subseteq \tdomain[P]$ of statements that can be, at least in line of principle, experimentally verified.\footnote{Yet, as we will see later, the precision of measurements will be quantified in terms of number of evolutions, not the range/uncertainty on the variables.}
-
-While we leave the full mathematical characterization in the mathematical section, there are a few things of note.
-Verifiable statements will correspond, in terms of sets of $E$, to finite and infinite volumes.
-Requiring those to be comparable means that $E$ is a measurable space and in phase space, for example, the standard measure will be recovered.
-As the process domain corresponds to a $\sigma$-algebra on $E$, the experimental domain corresponds to a topology on $E$.
-Therefore notions of continuity can be recovered as well.
-The process domain is the closure of the experimental domain under standard Boolean operations.
-This means that all the process domain is fully specify by, at least in principle, experimentally verifiable statements.
-The space does not contain any assertion that cannot ultimately be defined in terms of experiments.
-Lastly, requiring that, given enough time, each statement will be verified means the experimental domain must be generated by countably many statements.
-This means the topology is second countable and one is limited to work with ``nice'' spaces (e.g.
-spaces of continuous functions, separable Hilbert spaces, ...) not because they are ``nice'', but because they are the only physically meaningful ones.
-
-To sum up, we have two foundational object.
-The first object is the process domain $\tdomain[P]$ which is a Boolean algebra that contain all the possible description of our process.
-A subset $\edomain[P]$ contains those that are experimentally verifiable, which contain all the information in the full domain.
-The narrowest statements in $\tdomain[P]$ correspond to the evolutions $E$, that provide the full description at all times.
-The second object is the operator $\finer$ which gives us a way to compare two statements and decide which one gives a more refined description.
-Though two statements will not in general be comparable, all verifiable statements are comparable and all evolutions provide an equally refined description.
-
-This framework imposes very few requirements which are of very general applicability.
-It does not matter what specific quantities are used, or whether the process is from physics, chemistry, control systems, ecology, economics, sociology, or some other realm we have yet to discover.
-In the end, if we have a process we need to model, our theory will need at least to provide those objects.
-This means that the results we find will have broad applicability as well.
-
-\section{States}
-
-As the process extends over time, it is natural to want to organize the statements in our process domain into partial domains, one for each moment in time.
-Given our time parameter $t \in T \subseteq \mathbb{R}$, we can imagine to carve a subdomain $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
-Any logical operation on statements at fixed $t$ is also another statement at fixed $t$ therefore $\tdomain_t$ is a $\sigma$-complete Boolean sub-algebra of $\tdomain[P]$.
-The set $X_t \subset \tdomain_t$ of the narrowest statements will give the most precise description at that time: these are the possible states of our system.
-
-We should note that, technically, the statements are not defined at an infinitesimal moment $t$, rather within a finite interval $[t, t + \Delta t]$.
-This represent the time-scale at which the process will be described, meaning the description within $\tdomain_t$ should not be sensitive to what happens at faster scales.
-The quantities used to represent states at time $t$, then, should really be thought as averages within $\Delta t$.
-
-We should also note that there is nothing at this point requiring us that the set of descriptions $\tdomain_t$ and states $X_t$ will be the same at all times.
-
-What we want to do now is relate states at different times.
-In particular we want to formally define the notion of deterministic process: one in which, given the state at a given time, we can always predict the state at all future times.
-This can be done through the use of the narrowness operator: $\stmt_{t_1} \narrower \stmt_{t_2}$ means that all evolutions compatible with $\stmt_{t_1}$ are also compatible with $\stmt_{t_2}$.
-In other words, if we find $\stmt_{t_1}$ to be true we will also find $\stmt_{t_2}$ to be true as well.
-Therefore a process is deterministic if, for all $t_1, t_2 \in T$, given $x_{t_1} \in X_{t_1}$ we can find $x_{t_2} \in X_{t_2}$ such that $x_{t_1} \narrower x_{t_2}$.
-
-At each moment, we have a set of possible statements about the system corresponding to physically measurable properties.
-We call state the maximal description that can be given at a particular time.
-Given the state, we know all that can be said about the system in that moment.
-We define an evolution of the system as the maximal description of the system at all times.
-If we knew the exact evolution, we know all possible descriptions of the system at all times.
-Therefore picking an evolution is equivalent to picking a state at each time.
-A particular process will constrain how the system can evolve, it will constrain the possible evolutions.
-Therefore we formally define a process as a set of possible evolutions, the ones that are allowed by the process.
-
-As each evolution gives the total description of the system at all times, we assume that all evolutions give the same level of description.
-We also assume we have a way to compare sets of possible evolutions and determine which is bigger and which is smaller.
-Mathematically, we have a measure defined on the set of possible evolutions.
-Each description, if verified experimentally, will narrow the set of possible evolutions to the ones that are compatible with said description.
-Therefore each description, and in particular each state, can be assigned a value that corresponds to the number of evolutions compatible with it.
-We call the logarithm of the number of evolution the entropy associated to the description.
-
-We define a process to be deterministic if knowing the state at one time means we can predict with absolute certainty the state at a future time.
-If this is the case, all evolutions compatible with one state cannot split and must all pass through the same states at all times.
-During a deterministic process, then, the number of evolutions per state can only increase over time.
-The entropy associated to each state is a monotonically increasing function.
-
-Conversely, we define a process to be reversible if knowing the state at one time means we can reconstruct with absolute certainty the state a previous time.\footnote{Note that this notion of reversibility is not identical to the one usually found in thermodynamics, which corresponds to the ``ability to undo the change''.
-We will later find that the notion of thermodynamic reversibility corresponds to the case where system and environment together undergo a deterministic and reversible process.} Reasoning along the same lines, during a reversible process the entropy associated to each state can only decrease over time, as evolutions cannot merge.
-A deterministic and reversible process will necessarily see the number of states, and therefore the entropy, remain the same.
-
-In a deterministic process, a state will be an equilibrium if, roughly speaking, the evolutions have fully merged and they will all give the same future state.
-In other words, after the equilibria the evolution is effectively reversible.
-This means that, at equilibria, the entropy will have reached a maximum.
-Each equilibria will be associated with a bundle of evolutions which will define, at times before the equilibrium, the set of states that will merge.
-For each equilibria one can construct a constraint, an assertion that remains true for each equilibria.
-Therefore each equilibria can be seen as the maximizing the entropy under the constraints imposed by the process.
-
-Finally, a process over a system can be broken into two processes over two systems if and only if they are independent.
-That is, if the evolution over one subsystem tells us nothing about the evolution of the other or, equivalently, the measure on the evolutions on the whole is the product of the measure on the subsystems.
-In those cases, the entropy of the system is equal to the sum of the entropy of the subsystems.
-However, note that if the systems are independent throughout the whole evolution, no coupling can be defined between the subsystem.
-
-A useful scenario is when the systems can be considered independent at the beginning of the evolution, couple, and then return to be independent at equilibrium.
-The entropy can then be considered additive at the beginning and at the end of the evolution, but not in between.
-Equilibria on the whole will necessarily correspond to equilibria on each subsystem.
-Note how this conceptually corresponds to the scenarios studied in thermodynamics.
-
-NOTE: measure is invariant on Hamiltonian evolution, but not under irreversible evolution. Can't have the same measure at all times.
 

### 2020-09-01 Articles/2019-Entropy/Source/entropy-article.tex
Maybe done up to thermodynamics

@@
 
-As $\tdomain[P]$ contains different levels of description, one statement may give a more specific description than another.
+As $\tdomain[P]$ contains different levels of description, some statement will be more or less specific.
 For example, we say that \statement{the horizontal position is between 2.5 and 3 meters} is \textbf{narrower} than \statement{the horizontal position is between 2 and 3.5 meters}.
@@
 Every statement $\stmt \in \tdomain[P]$ can now be characterized by the set $A(\stmt) \subseteq E$ of all evolutions compatible with $\stmt$, that is all the evolutions for which $\stmt$ will be true.
-Mathematically,  $A(\edomain[P])$ is a topology over $E$ while $A(\tdomain[P])$ forms a $\sigma$-algebra over $E$, the Borel algebra of $A(\edomain[P])$.
+Mathematically,  $A(\edomain[P])$ is a \textbf{topology} over $E$ while $A(\tdomain[P])$ forms a $\sigma$-algebra over $E$, the \textbf{Borel algebra} of $A(\edomain[P])$.
 Narrowness and compatibility become set realtionships in the $\sigma$-algebra: $\stmt_1 \narrower \stmt_2$ if and only if $A(\stmt_1) \subseteq A(\stmt_2)$ (i.e.~all trajectories for which $\stmt_1$ is true are such that $\stmt_2$ is also true) and $\stmt_1 \comp \stmt_2$ if and only if $A(\stmt_1) \cap A(\stmt_2) \neq \emptyset$ (i.e.~two descriptions are compatible if there is an evolution that is compatbile with both).
+This connection gives a direct physical meaning to these foundational set-based mathematical structures.
 
@@
 
-As the process extends over time, we want to organize the statements in our process domain into partial domains, one for each moment in time.
+As a process extends over time and potentially describes multiple systems, we want to organize the statements into partial domains, one for each moment in time relative to a particular system.
 Given a \textbf{time parameter} $t \in T \subseteq \mathbb{R}$, we can imagine to carve a \textbf{system domain} $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
@@
 The set $X_t \subset \tdomain_t$ of the narrowest statements will give the most precise description at that time, we call these the  possible \textbf{snapshots} of our system at that time.
+A \textbf{trajectory} is a sequence $x(t)$ of snapshots at each time.
+A particular evolution will identify a trajectory: that is, given $\stmt[e] \in E$ we can find $x(t)$ such that $\stmt[e] \narrower x(t_0) \AND x(t_1) \AND x(t_2) \AND ...
+$.
+The converse is not true: a single trajectory may correspond to multiple evolutions, since the evolutions potentially describe the same system at greater accuracy and/or other systems.
 
-We should note that, technically, the statements are not defined at an infinitesimal moment $t$, rather within a finite interval $[t, t + \Delta t]$.
+In fact, we should note that, technically, the statements are not defined at an infinitesimal moment $t$, rather within a finite interval $[t, t + \Delta t]$.
 This represents the time-scale at which the process will be described, meaning the description within $\tdomain_t$ should not be sensitive to what happens at faster scales.
@@
 $A(x) \neq \emptyset$ for all $x \in X_t$), which may change in time.
-For example, the set of possible configurations for a system under dissaptive processes will shrink as they converge to equilibrium.
+For example, the set of possible configurations for a system under dissipative processes will shrink as they converge to equilibrium.
 In quantum mechanics, the possible configurations after a measurement are restricted to the eigenestates, and a different choice of measurement will lead to a different process with different possible evolutions.
@@
 In other words, the state space of the system (i.e.
-the set of all possible complete descriptions of the system and only about that system, independently of the process at hand) is something we'll construct later.
+the set of all possible complete descriptions of the system and only about that system, independently of the process at hand) is something we'll construct later.\footnote{In other words, starting directly from a state space comes with a set of implicit assumptions about the system which are at the foundations of thermodynamics specifically and physics more in general.}
 
-We say a process is \textbf{deterministic} over the system domains $\{X_t\}_{t \in T}$ if given the snapshot at one time we can always predict the snapshot at all future times.
-Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_1 \in X_{t_1}$ we can find an $x_2 \in X_{t_2}$ such that $x_1 \narrower x_2$.
-
+We say a process is \textbf{deterministic} over the system domains $\{\tdomain_t\}_{t \in T}$ if given the snapshot at one time we can always predict the snapshot at all future times.
+Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_1 \in X_{t_1} \subset \tdomain_{t_1}$ we can find an $x_2 \in X_{t_2} \subset \tdomain_{t_2}$ such that $x_1 \narrower x_2$.
 Intuitively, all the evolutions that pass through $x_1$ will also pass through $x_2$.
@@
 
-Conversely, we say a process is \textbf{reversible} over the system domains $\{X_t\}_{t \in T}$ if given the snapshot at one time we can always reconstruct the snapshot at all past times.
-Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_2 \in X_{t_2}$ we can find a $x_1 \in X_{t_1}$ such that $x_2 \narrower x_1$.
-
+Conversely, we say a process is \textbf{reversible} over the system domains $\{\tdomain_t\}_{t \in T}$ if given the snapshot at one time we can always reconstruct the snapshot at all past times.
+Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_2 \in X_{t_2} \subset \tdomain_{t_2}$ we can find a $x_1 \in X_{t_1} \subset \tdomain_{t_1}$ such that $x_2 \narrower x_1$.
 In this case, all the evolutions that pass through $x_2$ must have passed through $x_1$.
@@
 This is the core idea of the second law.
+
+These definitions naturally link to the idea of a dynamical system and clarify a critical issue.
+Suppose we have a deterministic process and we pick $t_1, t_2 \in T$.
+For each $x_1 \in X_1$, the final snapshot $x_2 \in X_2$ must be unique as all snapshots in $X_2$ are incompatible with each other.
+Therefore we can write a function $f : X_1 \to X_2$ such that $x_1 \narrower f(x_1) \equiv x_2$ that describes that particular step in the process.
+We call $f$ the \textbf{law of evolution}.
+Conversely, if the system is reversible, we would be able to find $g : X_2 \to X_1$ such that $x_2 \narrower g(x_2) \equiv x_1$.
+We call $g$ the \textbf{law of (inverse) evolution}.
+Therefore \textbf{a law of (possibly inverse) evolution can always be used to fully characterize deterministic and/or reversible processes}.
+
+We want to stress that if $f$ is invertible, the system is \emph{not} necessarily reversible.
+In fact, we'd have $x_1 \equiv f^{-1}(x_2) \narrower x_2$, which is not the same as $x_2 \narrower g(x_2) \equiv x_1$: narrowness is in the opposite direction.
+In other words, \textbf{an invertible deterministic system is not necessarily reversible}.
+For example, consider a damped harmonic oscillator: the dynamics is in principle invertible (we can reconstruct the trajectory) but it is not reversible (as the evolutions bunch together, finite precision knowledge of the initial conditions gives us more information of the same finite precision knowledge of the final conditions).
+These types of distinctions are crucial we'd like to stress how it emerges naturally within the framework.
 
@@
 
-Next step is to quantify the size of the sets of evolutions, specifically those that correspond to each statement.
-Since statements at a finer level of description will put a greater constrain to the set of evolutions, quantifying the size of $A(\stmt)$ means quantifying the granularity of the description of $\stmt$.
+Next step is to quantify the granularity of the description provided by each statement.
+Since statements at a finer level of description will put a greater constrain on the set of compatible evolutions, quantifying the granularity of the description of $\stmt$ means quantifying the size of $A(\stmt)$.
 The statement \statement{at time 0 s the particle is between 0 m and 1 m} is doubly more specific than \statement{at time 0 s the particle is between 0 m and 2 m} precisely because it correspond to half the evolutions of the second statement.
 
-The naive approach would be to assign a measure over $A(\tdomain[P])$: since this is is a $\sigma$-algebra over $E$, it would seem we are done.
+The naive approach would be to assign a measure over $A(\tdomain[P])$: since this is is a $\sigma$-algebra over $E$ and a measure is exactly one uses to give a size to each set, it would seem we are done.
 Unfortunately it is not that simple: a single measure can only compare objects of the same dimensionality.
@@
 
+The idea is that geometrical, measure theoretic, information theoretic and probabilistic structures all descend from this more fundamental structure, the partial order that describes this ``information granularity''.
+We note that all these structures can be used to determine a size and therefore a partial order on the set.
+It would be possible, given sufficient conditions, to recover those structure from the order itself.
+The mathematical detail of how this would exactly work (i.e.
+what condition are necessary and sufficient) is out of the scope of this work.
+Here we are interested in how the physical content attached to the partial order give rises to the other structure.
+
 To recover measures, we pick a statement $\stmt[u] \in \tdomain[P]$ and construct a \textbf{measure} $\mu_{\stmt[u]}$ such that $\mu_{\stmt[u]}(\stmt[u]) = 1$ and $\stmt_1 \finer \stmt_2$ will mean $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
@@
 If two statements are such that the measure of one with respect to the other is finite and non-zero, we say they are \textbf{finitely comparable}.
+We can show that if $\stmt[u]_1$ and $\stmt[u]_2$ are finitely comparable we have $ \mu_{\stmt[u]_1}(\stmt) = \mu_{\stmt[u]_1}(\stmt[u]_2) \mu_{\stmt[u]_2}(\stmt)$.
 Furthermore, since all evolutions give a complete description of the whole process, we will assume they are all equigranular.
@@
 If we consider all the evolutions compatible with \statement{the coin is fair}, we must have that half of them are compatible with \statement{the coin is head} while the other are compatible with \statement{the coin is tails}.
-Therefore we have that the \textbf{probability} $P(\stmt_2 | \stmt_1) = \mu_{\stmt_1}(\stmt_1 \AND \stmt_2)$ where $\stmt_1 \AND \stmt_2$ is the logical conjunction (AND) between the two statements.
+Therefore we define that the \textbf{probability} of $\stmt_2$ given $\stmt_1$ is
+\begin{equation}
+	P(\stmt_2 | \stmt_1) = \mu_{\stmt_1}(\stmt_1 \AND \stmt_2) = \frac{\mu_{\stmt[u]}(\stmt_1 \AND \stmt_2)}{\mu_{\stmt[u]}(\stmt_1)}
+\end{equation} where $\stmt_1 \AND \stmt_2$ is the logical conjunction (AND) between the two statements and $\stmt[u]$ is finitely comparable to $\stmt_1$.
 In other words, the probability quantifies the ratio of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$.
+This can be shown to satisfy the standard axioms of probability.
 
-We can study a one-step process by tracking the probability $P(\stmt_1 | \stmt_0)$ that the final description $\stmt_1$ is true given that the initial description $\stmt_0$ is true, which is simply the fractions of evolutions starting in $\stmt_0$ (i.e.
-are compatible to $\stmt_0$)) that end in $\stmt_1$.
-The fact that \textbf{probability theory can be used to fully characterize all one-step processes}, the simplest processes, tells us why it is so important in physics.
+Suppose that, given a process, we pick $t_1, t_2 \in T$.
+Given an initial description $\stmt_1 \in \tdomain_{t_1}$ and a final description $\stmt_2 \in \tdomain_{t_2}$, the \textbf{forward transition probability} $P(\stmt_2 | \stmt_1)$ tell us what is the fraction of evolutions starting in $\stmt_1$ (i.e.
+are compatible to $\stmt_1$)) that end in $\stmt_2$.
+In the same way, we can look at the \textbf{backwards transition probability}.
+The probability space is enough to recover the process, which means \textbf{forward and backward transition probabilities can always be used to fully characterize a process.}
+
+We say that a process is \textbf{isoentropic} if for any possible trajectory $x(t)$ we have $S_{\stmt[u]}(x(t)) = S_{\stmt[u]}(x(t + \Delta t)) \neq 0$.
+We saw that all deterministic and reversible processes are isoentropic, but the converse is not true.
+We can give, though, a characterization in terms of transition probabilities.
+Assume $\stmt[u]$, $x(t)$ and $x(t+\Delta t)$ are all finitely comparable.
+We can write:
+\begin{align*}
+	\mu_{\stmt[u]}(x(t) &\AND x(t+\Delta t)) \\
+	&=  \mu_{\stmt[u]}(x(t)) P(x(t)|x(t+\Delta t)) \\
+	&=  \mu_{\stmt[u]}(x(t+\Delta t)) P(x(t+\Delta t)|x(t)) \\
+\end{align*}
+Therefore
+\begin{equation}
+\frac{\mu_{\stmt[u]}(x(t))}{\mu_{\stmt[u]}(x(t+\Delta t))} = \frac{P(x(t+\Delta t)|x(t))}{P(x(t)|x(t+\Delta t))}.
+\end{equation}
+If the forward and backward probabilities are the same, the initial and final snapshots have the same measure.
+Conversely, if the initial and final have the same measure, the forward and backward probabilities must be the same.
+Equal measure corresponds to equal process entropy.
+Therefore \textbf{isoentropic processes are those and only those for which backward and forward probabilities are the same}.
 
@@
 
-So far we have made no guarantees as to whether each $\tdomain_t$ contains descriptions that refer only to a particular physical object.
-So we need to understand what is needed to be able to define a system and its state space.
-Intuitively, the state space should provide us with all the descriptions of the system, and only of the system, at any point in time for any process.
-We are therefore asked to create an abstraction that stitches together all possible descriptions, all possible snapshots, about the system and only about the system, which means there are a few issues we need to consider.
+So far we have made no guarantees as to whether each domain $\tdomain_t$ and its snapshots $X_t$ contains \emph{all} descriptions that refer \emph{only} to a particular physical object.
+Some snapshots may or may not be available at a particular time and/or there may be correlations with other systems.
+We want the state space to consist of all the possible descriptions of a system, and only of that system, at any point in time for any process.
+This desire comes with some hidden requirements that we want to bring to light.
 
@@
 The process has to be such that the system is recognizable to be the same, our descriptions must still make sense.
-We can imagine a \textbf{state domain} $\tdomain$ of all statement fragments like \statement{the position of the ball is between 2 and 3 meters} together with a surjective map $\iota : \tdomain \times T \twoheadrightarrow \tdomain_t$ that adds \statement{at time t} for each possible time.
+Mathematically, we have a \textbf{state domain} $\tdomain$ of all statement fragments like \statement{the position of the ball is between 2 and 3 meters} together with a surjective map $\iota : \tdomain \times T \twoheadrightarrow \tdomain_t$ that adds \statement{at time t} for each possible time.
 The map $\iota$ must be a surjection that preserves the logical structure (i.e.
@@
 Since we can only properly talk about a system when it is independent from the other, what can we do in the other cases?
-The idea that we can still talk about a system in that case if we are able, at least in line of principle, to decouple the system from all others while not changing the nature of the system itself.
+The idea is that we can still talk about a system in that case if we are able, at least in line of principle, to decouple the system from all others while not changing the nature of the system itself.
 For example, we can talk about a molecule of a gas because we have a way to take such molecule out of the gas.
@@
 \end{equation}
-The process entropy increases and is maximized when the system is independent.\footnote{Mathematically, the relaxation process in needed to compare the process entropies for the correlated descriptions.
-As the same description can be more or less correlated in different cases, the independent ones work as a reference and we need a way to relate to them.}
-
-Note that all snapshots that relax to the same state will have a process entropy lower or equal to the state entropy.
-Therefore \textbf{the state entropy is the largest possible process entropy} for which the description given by the state applies.
-This does not simply represent a subjective information theoretic consideration (e.g.
+This means that all snapshots that relax to the same state will have a process entropy lower or equal to the state entropy.
+Therefore \textbf{the state entropy is the largest possible process entropy for which the description given by the state applies}.\footnote{Mathematically, the relaxation process in needed to compare the process entropies for the correlated descriptions.
+As the same description can be more or less correlated in different cases, the independent ones work as a reference and we need a way to relate to them.} This does not simply represent a subjective information theoretic consideration (e.g.
 we know only about the system); it is a physical consideration (e.g.
 knowledge about one system tells us nothing about the others).
-As we make the system more independent, there are more possible evolutions the combined system can have.
-Therefore in this framework \textbf{entropy maximization is equivalent to requiring system independence}.
-Entropy maximization, more than a law of physics, is a requirement to properly define the state space of a system, a requirement for doing physics.
-As such, we are therefore biased towards looking for such situations.
+As we make the system more independent, there are more possible evolutions the combined system can have, the dynamics of the system is less affected, less correlated, by the dynamics of the other.
+
+The takeaway message is that \textbf{entropy maximization is equivalent to requiring system independence}.
+Maximizing entropy means taking the system apart from the others, so that its dynamics and description is decoupled.
+Entropy maximization, more than a law of physics, is a requirement to properly define a part of nature as a system.
+When we divide nature in independent systems, therefore, we are automatically looking for processes that increase entropy.
 

### 2020-09-03 Articles/2019-Entropy/Source/entropy-article.tex
Recovering 3 laws

@@
 
+If the final state is identified by the set of state variables $\xi^a : X \to \mathbb{R}$, the initial snapshot will be identified by the same quantities since the relaxation process should not, in principle, disturb the system.
+These quantities, then, are a constraint during relaxation and, additionally, they must be the only constraints.
+Additional constraints, in fact, would increase the number of final states, meaning the initial set of quantities did not form a full set of state variables.
+Therefore \textbf{state variables identify both the states of the systems and the constraints during relaxation}.
+
 The takeaway message is that \textbf{entropy maximization is equivalent to requiring system independence}.
 Maximizing entropy means taking the system apart from the others, so that its dynamics and description is decoupled.
+We should stress here that the description may not include all possible degrees of freedoms of the object studied (e.g.
+macrostate variable, position and momentum of the center of mass).
+Maximizing entropy means requiring independence of that partial description from the rest, effectively decoupling environment, other systems and internal dynamics in one operation.
 Entropy maximization, more than a law of physics, is a requirement to properly define a part of nature as a system.
@@
 
+As we have seen, the definition of a system as a part of nature to be studied independently from the rest comes with physical requirements that provide a rich conceptual and mathematical structure that maps extremely well to the basic ideas in thermodynamics.
+
 \subsection{Thermodynamic processes}
+
+We now have all the basic tools we need to recover thermodynamic as the study of the relaxation process of composite systems.
+That is, we have two systems, we put them together and relax the composite system.
+In light of what discussed before, the composition rule must tell us what composite states are possible and what is the relaxation process.
+These will be linked because of the double role of the state variables.
+Thermodynamic describes that link in the case of extensive (i.e.
+additive) state variables.\footnote{It should be stressed that, in our framework, recovering the laws of thermodynamics does not always mean recovering universal truths about nature.
+It simply means identify the necessary and sufficient assumptions for which a given system will be suitable described by those laws.}
+
+As we noted before, given the same physical object, we will have a different relaxation process and different state variables.
+When composing two systems, we are automatically given a relaxation process which is the independent relaxation of each system.
+The state space, in this case, is simply the set of all possible pairs of states of the components.
+This means we have constraints within the system which is therefore not necessarily relaxed: the description of the whole is tied to the description of the parts.
+That is, we don't simply describe what the total energy or the total number of particle is, we also describe the energy and the particles for each part.
+We now want to remove the constraints within the system and characterize the relaxation process.
+We call \textbf{equilibrium} the final states of a relaxation process of a composite system with no internal constraints.
+We call one-step equilibration process a one step process that ends in equilibrium.
+
+\textbf{The first assumption in thermodynamic is that an equilibrium state of the whole is described by a pair of equilibrium states of the parts.} We qualify this as an assumptions as it not always true for all systems and can fail in several ways.
+For example, removing the wall between two substances may change them both in a third (this usually addressed by changing the state space of the parts to accommodate).
+More subtly, in the presence of impermeable walls we can characterize the state on either side with a fixed number of particle.
+When the wall is removed, the state will be characterized by an average number of particles.
+The state space is not technically the same.
+If the boundary effects between the two components are non negligible, they will not be independent (i.e.
+the entropy does not sum).
+Outside of thermodynamics, an entangled state is certainly a state of a composite system (e.g.
+two photons) where constraints are put only on the whole (e.g.
+total spin) and it cannot be described as a simple combination of subsystems' states (e.g.
+is it not separable).
+Therefore there are definitely cases that do not follow this assumption.
+Characterizing all cases in which the assumptions hold (e.g.
+a ``large'' system in contact with a sufficiently ``large reservoir'') is left for future work.
+
+\textbf{The second assumption in thermodynamic is that constraints are independent, each identified by an extensive (i.e.
+additive under composition) quantity.} There is good reason to suspect that this assumption is, at least in large part, already contained in the previous one.
+Under the first assumption, in fact, state entropy is linear under composition.
+It could be that, assuming the constraints are independent (i.e.
+the variation on one does not change the others) is enough to construct a linear quantity to describe them.
+In the end, a linear scale for energy, mass, volume, number of particles, etc..
+is typically construct by preparing reference sample and combining them.
+We leave this investigation for future work.
+
+\textbf{The third assumption in thermodynamic is that one such constraint is the internal energy and its corresponding equilibrium is thermal equilibrium.} This is hope that this too can be already contained in the first assumption.
+In previous work we have recovered both classical Hamiltonian and quantum unitary evolution under the assumption of deterministic and reversible evolution which, as we have seen, is equivalent to conservation of process entropy.
+In those settings the notion of energy, in the form of the Hamiltonian, arises naturally so it could be made to arise naturally here as well.
+Yet, there some technical subtleties in the relationship between the Hamiltonian and the internal energy which we still have to work out.
+
+Under these assumptions, we can recover the laws of thermodynamics. Let $A$ and $B$ be two system in thermal equilibrium and let $AB$ be the composite. We have:
+\begin{equation}
+	\begin{aligned}
+	U_{AB} &= U_A + U_B \\
+	S_{AB}(U_{AB}) &= S_A(U_A) + S_B(U_B) \\
+	&= \sup \{S_A(x) + S_B(y) \, | \, x+y= U_{AB}\} \\
+	\frac{1}{kT_A} = \beta_A = \frac{\partial S_A}{\partial U_A} &= \frac{\partial S_B}{\partial U_B} = \beta_B = \frac{1}{kT_B}
+	\end{aligned}
+\end{equation}
+which follows the standard textbooks derivations.
+One finds that thermal equilibrium is achieved if and only if $T_A$ and $T_B$ are equal.
+If $B$ is in thermal equilibrium with another system $C$ we also have $T_B$ is equal to $T_C$ and therefore $A$ and $C$ are also in thermal equilibrium.
+This recovers the zeroth law.
+
+We say a system $R$ is a \textbf{thermal reservoir} if its equation of state can be sufficiently approximated by $S = \beta U = \frac{U}{k T}$.
+Internal energy is the only relevant constraint for the thermal reservoir and the only way the entropy of the system can change.
+We call \textbf{heat} the energy exchanged by a thermal reservoir and we note $Q = - \Delta U$ as the energy lost by the reservoir during a one-step process.
+
+Let $M$ be a purely mechanical system. Let $U$ be the internal energy and $X^i$ the other variables. We have:
+\begin{equation}
+\begin{aligned}
+S_M&=S_M(W, X^i)=0 \\
+dS_M = 0 &= \frac{\partial S_M}{\partial U} dU + \frac{\partial S_M}{\partial X^i} dX^i \\
+dW &= \frac{1}{\frac{\partial S_M}{\partial U}}\frac{\partial S_M}{\partial X^i} dX^i \\
+P_i &= \frac{1}{\frac{\partial S_M}{\partial U}}\frac{\partial S_M}{\partial X^i} \\
+dU &= P_i dX^i.
+\end{aligned}
+\end{equation}
+For a purely mechanical system, a chance in internal energy is always linked to a change one of the other constraints.
+The energy of the purely mechanical system is \textbf{stored} in one of the constraints.
+We note a change of internal energy during a one-step equilibration process of a pure mechanical system as $W=\Delta U$.
+
+We now consider a system composed of a generic system $A$, a thermal reservoir $R$ and a purely mechanical system $M$.
+During a one-step equilibration process, because the total energy of the system must be conserved, we have:
+\begin{equation}
+\begin{aligned}
+&\Delta U_A + \Delta U_R + \Delta U_M = 0 \\
+&\Delta U_A - Q + W = 0 \\
+&\Delta U_A = Q - W.
+\end{aligned}
+\end{equation}
+This recovers the first law.
+
+We now look at the change in entropy, which can only increase during a one-step process. Keeping in mind $M$ is purely mechanical, we have:
+\begin{equation}
+\begin{aligned}
+&\Delta S_A + \Delta S_R + \Delta S_M \geq 0 \\
+&\Delta S_A \geq - \Delta S_R \\
+&\Delta S_A \geq \int \beta_R dQ \\
+&\Delta S_A \geq \int \frac{dQ}{k T_R}.
+\end{aligned}
+\end{equation}
+which recovers the second law.
+
+As for the second law, we have given a characterization that is far more general than what it is given in thermodynamics. Nevertheless, we can
 
@@
 We say a system is in \textbf{equilibrium} if the state tells us nothing about the
+
+ The physical meaning that ``all heat is equivalent'' and ``all diathermal walls are equivalent'' is explicit in the fact that the internal energy identifies the constraint.
 

### 2020-09-06 Articles/2019-Entropy/Source/entropy-article.tex
Review with Christine

@@
 
-The overall goal is to describe a generic system as it evolves in time by defining a minimal set of concepts that such a setup must define, and understand the tacit assumptions underlying state spaces and evolution laws.
-We want to find notions that can be applied in different settings, being it thermodyamics and/or dynamical systems alike.
-Then we want to derive general results based on such definitions.
-Here we provide an overview of the main features, leaving the formal definitions in the (yet to be written) appendix.
+The overall goal is to describe a generic system as it evolves in time by defining a minimal set of concepts that such a setup must define, thus bringing to light the tacit assumptions underlying state spaces and evolution laws.
+We want to find notions that are equally valid and useful within different settings (e.g.
+thermodynamics, dynamical systems, classical mechanics, ...).
+We provide here an overview of the main features and a firm idea of how they recover the more standard results, leaving a complete formal development to future works.
 
@@
 Our starting point is the \textbf{process domain} $\tdomain[P]$, which is the set that contains all possible descriptions of the system at all levels of detail at all possible times.
-Depending on the process, it will contain statements like \statement{the average volume between 1 and 2 seconds is between 3 and 4 liters} or \statement{the trajectory of the particle is $y=10 \, m - t^2 \cdot 9.80665 \, m/s^2 $}.
+Depending on the process, it will contain statements like \statement{the average volume of the system between 1 and 2 seconds is between 3 and 4 liters} or \statement{the trajectory of the particle is $y=10 \, m - t^2 \cdot 9.80665 \, m/s^2 $}.
 It does not matter at this point what the statements themselves are, except they must be, at least in line of principle, defined from experimentally well-defined starting points.
+The translation of these physical requirements into a formal structure has already been carried out in previous work, and we will briefly summarize them in this section.
 Mathematically, $\tdomain[P]$ is a $\sigma$-complete Boolean algebra\footnote{A set of statements that is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).} and it is the closure of the set of experimentally verifiable statements $\edomain[P] \subseteq \tdomain[P]$.\footnote{The set $\edomain[P]$ will form a Heyting algebra which, to be experimentally reachable, must allow a countable basis.
@@
 
-As $\tdomain[P]$ contains different levels of description, some statement will be more or less specific.
+As $\tdomain[P]$ contains different levels of description, some statements will be more or less specific.
 For example, we say that \statement{the horizontal position is between 2.5 and 3 meters} is \textbf{narrower} than \statement{the horizontal position is between 2 and 3.5 meters}.
 Note that whenever the first statement is true, the second one must be true as well.
-Given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$, we formally capture this relationship by noting $\stmt_1 \narrower \stmt_2$.
+Given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$, we formally capture this relationship by noting $\stmt_1 \narrower \stmt_2$ (read ``is narrower than'').
 Narrowness can describe relationships at different scales, on different quantities (e.g.
 \statement{the horizontal position is between 2.5 and 3 meters and the vertical position is between 1 and 1.5 meters} $\narrower$ \statement{the horizontal position is between 2.5 and 3 meters}), constraints between variables at the same time (e.g.
-\statement{the temperature of the water in the glass is 3.98 C} $\narrower$ \statement{the density of the water in the glass is 1 $g/cm^3$}) or at different times (e.g.
+\statement{the temperature of the water in the glass is 3.98 $^\circ C$} $\narrower$ \statement{the density of the water in the glass is 1 $g/cm^3$}) or at different times (e.g.
 \statement{at time 0 s the position is 1 m and the velocity is 1 m/s} $\narrower$ \statement{at time 1 s the position is 2 m}).
-It is a general tool to capture relationships within the theory which intuitively can also be thought as implication.\footnote{Technically, implication in classical logic is a truth function as classical logic does not incorporate the idea of different ``possible cases''.
+It is a general tool to capture relationships within the theory which intuitively can also be thought of as implication.\footnote{Technically, implication in classical logic is a truth function as classical logic does not incorporate the idea of different ``possible cases''.
 What we have is more similar in spirit to semantic consequence in modal logic, though modal logic brings in a lot of undesirable elements we do not want.}
@@
 For example, we want to say that \statement{the horizontal position is between 2.5 and 3.5 meters} is \textbf{compatible} with \statement{the horizontal position is between 2 and 3 meters}.
-In general, given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$ we note $\stmt_1 \comp \stmt_2$ if it is possible for them to be both true.
+In general, given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$ we denote $\stmt_1 \comp \stmt_2$ (read ``is compatible with'') if it is possible for them to be both true.
 As with narrowness, compatibility can be in terms of different ranges of the same quantities, of different quantities and of constraints at equal or different times.
 
-Finally, we want to capture whether the truth of one or more statements incluences the truth of others.
-For example, we want to say that \statement{the temperature of the water in the glass is 3.98 C} is \textbf{independent} from \statement{the volume of the water in the glass is 10 $cL$} but is not independent from \statement{the density of the water in the glass is 1 $g/cm^3$}.
-We note $\stmt_1 \indep \stmt_2$ when two statements are independent.
+Finally, we want to capture whether the truth of one or more statements influences the truth of others.
+For example, we want to say that \statement{the temperature of the water in the glass is 3.98 $^\circ C$} is \textbf{independent} from \statement{the volume of the water in the glass is 10 $cL$} but is not independent from \statement{the density of the water in the glass is 1 $g/cm^3$}.
+We note $\stmt_1 \indep \stmt_2$ (read ``is independent of'') when two statements are independent.
 
@@
 Mathematically,  $A(\edomain[P])$ is a \textbf{topology} over $E$ while $A(\tdomain[P])$ forms a $\sigma$-algebra over $E$, the \textbf{Borel algebra} of $A(\edomain[P])$.
-Narrowness and compatibility become set realtionships in the $\sigma$-algebra: $\stmt_1 \narrower \stmt_2$ if and only if $A(\stmt_1) \subseteq A(\stmt_2)$ (i.e.~all trajectories for which $\stmt_1$ is true are such that $\stmt_2$ is also true) and $\stmt_1 \comp \stmt_2$ if and only if $A(\stmt_1) \cap A(\stmt_2) \neq \emptyset$ (i.e.~two descriptions are compatible if there is an evolution that is compatbile with both).
+Narrowness and compatibility become set relationships in the $\sigma$-algebra: $\stmt_1 \narrower \stmt_2$ if and only if $A(\stmt_1) \subseteq A(\stmt_2)$ (i.e.~all evolutions for which $\stmt_1$ is true are such that $\stmt_2$ is also true) and $\stmt_1 \comp \stmt_2$ if and only if $A(\stmt_1) \cap A(\stmt_2) \neq \emptyset$ (i.e.~two descriptions are compatible if there is an evolution that is compatible with both).
 This connection gives a direct physical meaning to these foundational set-based mathematical structures.
@@
 	\hline 
-	$\indep$ & independent & Two statements are independent  $\stmt_1 \comp \stmt_2$ if the truth of one does not influence the truth of the other \\ 
+	$\indep$ & independent & Two statements are independent  $\stmt_1 \indep \stmt_2$ if the truth of one does not influence the truth of the other \\ 
 \hline 
@@
 As a process extends over time and potentially describes multiple systems, we want to organize the statements into partial domains, one for each moment in time relative to a particular system.
-Given a \textbf{time parameter} $t \in T \subseteq \mathbb{R}$, we can imagine to carve a \textbf{system domain} $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
-In particular, we define a \textbf{one-step processes} one that only has an \textbf{initial domain} $\tdomain_{t_0}$ and a \textbf{final domain} $\tdomain_{t_1}$.
+Given a \textbf{time parameter} $t \in T \subseteq \mathbb{R}$, we can imagine carving out a \textbf{system domain} $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time.
+In particular, we define \textbf{one-step process} as one that only has an \textbf{initial domain} $\tdomain_{t_0}$ and a \textbf{final domain} $\tdomain_{t_1}$.
 Note that $\tdomain_t$ is a $\sigma$-complete Boolean sub-algebra of $\tdomain[P]$ since any logical operation on statements at fixed $t$ is also another statement at fixed $t$.
-The set $X_t \subset \tdomain_t$ of the narrowest statements will give the most precise description at that time, we call these the  possible \textbf{snapshots} of our system at that time.
+The set $X_t \subset \tdomain_t$ of the narrowest statements will give the most precise description at that time; we call these the  possible \textbf{snapshots} of our system at that time.
 A \textbf{trajectory} is a sequence $x(t)$ of snapshots at each time.
@@
 
-In fact, we should note that, technically, the statements are not defined at an infinitesimal moment $t$, rather within a finite interval $[t, t + \Delta t]$.
+In fact, we should note that, technically, the statements are not defined at an infinitesimal moment $t$, but rather within a finite interval $[t, t + \Delta t]$.
 This represents the time-scale at which the process will be described, meaning the description within $\tdomain_t$ should not be sensitive to what happens at faster scales.
-The quantities used to represent states at time $t$, then, should really be thought as averages within $\Delta t$.
+The quantities used to represent states at time $t$, then, should really be thought of as averages within $\Delta t$.
 In the same way, $\tdomain_t$ will in general represent the system at the chosen level of description, not the narrowest level possible.
@@
 We should also note that the set of possible snapshots $X_t$ may not be the same at all times and may not correspond to the state of the system.
-First of all $X_t$ will only include those configurations that are compatible with at least by one possible evolution (i.e.
+First of all $X_t$ will only include those configurations that are compatible with at least one possible evolution (i.e.
 $A(x) \neq \emptyset$ for all $x \in X_t$), which may change in time.
 For example, the set of possible configurations for a system under dissipative processes will shrink as they converge to equilibrium.
-In quantum mechanics, the possible configurations after a measurement are restricted to the eigenestates, and a different choice of measurement will lead to a different process with different possible evolutions.
+In quantum mechanics, the possible configurations after a measurement are restricted to the eigenstates, and a different choice of measurement will lead to a different process with different possible evolutions.
 Secondly, the best possible description may be broader than the full state of the system (e.g.
@@
 Conversely, we say a process is \textbf{reversible} over the system domains $\{\tdomain_t\}_{t \in T}$ if given the snapshot at one time we can always reconstruct the snapshot at all past times.
-Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_2 \in X_{t_2} \subset \tdomain_{t_2}$ we can find a $x_1 \in X_{t_1} \subset \tdomain_{t_1}$ such that $x_2 \narrower x_1$.
+Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_2 \in X_{t_2} \subset \tdomain_{t_2}$ we can find an $x_1 \in X_{t_1} \subset \tdomain_{t_1}$ such that $x_2 \narrower x_1$.
 In this case, all the evolutions that pass through $x_2$ must have passed through $x_1$.
@@
 If we limit ourselves to deterministic processes, with these few definitions, we have already found that the set of possible evolutions ``grows bigger'' under irreversible evolution while it remains the ``the same size'' under reversible evolution.
-This is the core idea of the second law.
+This is the core idea that will recover the second law of thermodynamics.
 
@@
 We want to stress that if $f$ is invertible, the system is \emph{not} necessarily reversible.
-In fact, we'd have $x_1 \equiv f^{-1}(x_2) \narrower x_2$, which is not the same as $x_2 \narrower g(x_2) \equiv x_1$: narrowness is in the opposite direction.
+In fact, we would have $x_1 \equiv f^{-1}(x_2) \narrower x_2$, which is not the same as $x_2 \narrower g(x_2) \equiv x_1$: narrowness is in the opposite direction.
 In other words, \textbf{an invertible deterministic system is not necessarily reversible}.
-For example, consider a damped harmonic oscillator: the dynamics is in principle invertible (we can reconstruct the trajectory) but it is not reversible (as the evolutions bunch together, finite precision knowledge of the initial conditions gives us more information of the same finite precision knowledge of the final conditions).
-These types of distinctions are crucial we'd like to stress how it emerges naturally within the framework.
+For example, consider a damped harmonic oscillator: the dynamics is in principle invertible (we can reconstruct the trajectory) but it is not reversible (as the evolutions bunch together, finite precision knowledge of the initial conditions gives us more information than the same finite precision knowledge of the final conditions).
+These types of distinctions are crucial and they emerge naturally within the framework.
 
@@
 
-Next step is to quantify the granularity of the description provided by each statement.
-Since statements at a finer level of description will put a greater constrain on the set of compatible evolutions, quantifying the granularity of the description of $\stmt$ means quantifying the size of $A(\stmt)$.
-The statement \statement{at time 0 s the particle is between 0 m and 1 m} is doubly more specific than \statement{at time 0 s the particle is between 0 m and 2 m} precisely because it correspond to half the evolutions of the second statement.
+The next step is to quantify the granularity of the description provided by each statement.
+Since statements at a finer level of description will put a greater constraint on the set of compatible evolutions, quantifying the granularity of the description of $\stmt$ means quantifying the size of $A(\stmt)$.
+The statement \statement{at time 0 s the particle is between 0 m and 1 m} is doubly more specific than \statement{at time 0 s the particle is between 0 m and 2 m} precisely because it corresponds to half the evolutions of the second statement.
 
-The naive approach would be to assign a measure over $A(\tdomain[P])$: since this is is a $\sigma$-algebra over $E$ and a measure is exactly one uses to give a size to each set, it would seem we are done.
+The naive approach would be to assign a measure over $A(\tdomain[P])$: since this is a $\sigma$-algebra over $E$ and a measure is exactly what one uses to give a size to each set, it would seem we are done.
 Unfortunately it is not that simple: a single measure can only compare objects of the same dimensionality.
-If we imagine all evolutions to be points on an $n$ dimensional manifold, if the measure gives finite values for $k$-dimensional regions it will necessarily give measure zero for all regions with lesser dimensionality and infinite measure for those with greater dimensionality.
-Coarse grained and subsystem descriptions will have different dimensionality as they describe different numbers of degrees of freedom, therefore we need a way to quantify objects at all levels.\footnote{Typically one uses geometrical structure for his purpose.
+If we imagine all evolutions to be points on an $n$-dimensional manifold, if the measure gives finite values for $k$-dimensional regions it will necessarily give measure zero for all regions with lesser dimensionality and infinite measure for those with greater dimensionality.
+Coarse grained and subsystem descriptions will have different dimensionality as they describe different numbers of degrees of freedom, therefore we need a way to quantify objects at all levels.\footnote{Typically one uses geometrical structures for his purpose.
 In phase space, the symplectic form allows us to define areas of two-dimensional surfaces, and use those to define volumes or even-dimensional areas.
@@
 
-Therefore our starting point will be the ability to compare two statements and decide which one gives a refined description of the process.
-Formally, we have partial order $\finer$ on $\tdomain[P]$ which, given two statements, it tells us whether the description of one is \textbf{finer}, more refined, than the other.
+Therefore our starting point will be the ability to compare two statements and decide which one gives a finer description of the process.
+Formally, we have partial order $\finer$ on $\tdomain[P]$ which, given two statements, tells us whether the description of one is \textbf{finer}, more refined, than the other.
 Two statements are \textbf{equigranualar}, noted $\stmt_1 \eqgran \stmt_2$, if they provide the same level of description.
@@
 We say two statements are \textbf{comparable} if one is finer than the other.
-The order is partial because not all statements are comparable to each other.
+The order $\finer$ is partial because not all statements are comparable to each other.
 For example, consider the pressure/volume state space for an ideal gas.
-Comparing \statement{the pressure is 1 kPa and the volume is between 1 and 2 liters} and \statement{the pressure is between 1 and 2 kPa and the volume is 1 liter} would mean saying where one unit liter is bigger or smaller than 1 kPa.\footnote{In phase space, for example, we cannot in general compare $q^i$ and $p_i$, yet we can always compare areas of phase space.}
+Comparing \statement{the pressure is 1 kPa and the volume is between 1 and 2 liters} and \statement{the pressure is between 1 and 2 kPa and the volume is 1 liter} would mean saying whether one liter is bigger or smaller than one kPa.\footnote{In phase space, for example, we cannot in general compare $q^i$ and $p_i$, yet we can always compare areas of phase space} The idea of physical dimension would therefore be something that is formally captured mathematically.
 
@@
 We note that all these structures can be used to determine a size and therefore a partial order on the set.
-It would be possible, given sufficient conditions, to recover those structure from the order itself.
-The mathematical detail of how this would exactly work (i.e.
-what condition are necessary and sufficient) is out of the scope of this work.
-Here we are interested in how the physical content attached to the partial order give rises to the other structure.
+It would be possible, given sufficient conditions, to recover those structures from the order itself.
+The mathematical detail of how exactly this would work (i.e.
+what conditions are necessary and sufficient) is out of the scope of this work.
+Here we are interested in how the physical content attached to the partial order gives rise to the other structures.
 
-To recover measures, we pick a statement $\stmt[u] \in \tdomain[P]$ and construct a \textbf{measure} $\mu_{\stmt[u]}$ such that $\mu_{\stmt[u]}(\stmt[u]) = 1$ and $\stmt_1 \finer \stmt_2$ will mean $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
-That is, we can quantity how more or less precise a statement is compared to a fixed statement taken as a unit.
+To recover measures, we pick a statement $\stmt[u] \in \tdomain[P]$, called unit, and construct a \textbf{measure} $\mu_{\stmt[u]}$ such that $\mu_{\stmt[u]}(\stmt[u]) = 1$ and $\stmt_1 \finer \stmt_2$ will mean $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
+That is, we can quantify how much more or less precise a statement is compared to a fixed statement taken as a unit.
 If two statements are such that the measure of one with respect to the other is finite and non-zero, we say they are \textbf{finitely comparable}.
@@
 Furthermore, since all evolutions give a complete description of the whole process, we will assume they are all equigranular.
-Therefore if we pick $\stmt[u] \in E$, then $\mu_{\stmt[u]}(\stmt)=\#(A(\stmt))$ would be the counting measure of evolutions compatible with $\stmt$.\footnote{We still needs to understand the necessary and sufficient conditions under which such measures would exist and be unique.
-For the present work, we assume it can be done.
-We know that it is a necessary condition that all evolution must be equigranular.
-It remains an open question is whether it is sufficient.}
+Therefore if we pick $\stmt[u] \in E$, then $\mu_{\stmt[u]}(\stmt)=\#(A(\stmt))$ would be the counting measure of evolutions compatible with $\stmt$.\footnote{We still need to understand the necessary and sufficient conditions under which such measures would exist and be unique.
+For the present work, we assume it can be done.}
 
-We can also define the \textbf{evolution entropy} with respect to $\stmt[u]$ the quantity $S_{\stmt[u]}(\stmt) = \log \mu_{\stmt[u]}(\stmt)$.
-This can be thought as the number of bits, yes/no questions, that separate the level of description of $\stmt[u]$ from the one provided by $\stmt$.
+We can also define the \textbf{evolution entropy} with respect to $\stmt[u]$ as the quantity $S_{\stmt[u]}(\stmt) = \log \mu_{\stmt[u]}(\stmt)$.
+This can be thought of as the number of bits, yes/no questions, that separate the level of description of $\stmt[u]$ from the one provided by $\stmt$.
 Connecting the definitions from above, under a deterministic process we have $S_{\stmt[u]}(x(t)) \leq S_{\stmt[u]}(x(t + \Delta t))$ since $x(t)\narrower x(t+\Delta t)$ means $x(t) \finer x(t+\Delta t)$ and $\mu_{\stmt[u]}(x(t)) \leq \mu_{\stmt[u]}(x(t+\Delta t))$.
@@
 
-Quantifying evolutions allows us to define probability as well.
-If we consider all the evolutions compatible with \statement{the coin is fair}, we must have that half of them are compatible with \statement{the coin is head} while the other are compatible with \statement{the coin is tails}.
+Quantifying evolutions allows us to define probability.
+The idea is a process can be realized differently by each possible evolution.
+The more evolutions a statement is compatible with, the more it is likely to happen.
+For example, consider a process that implements a coin flip.
+If the coin (and the process flipping the coin) is fair, we must have that half the evolutions are compatible with \statement{the coin is head} while the other half are compatible with \statement{the coin is tails}.
 Therefore we define that the \textbf{probability} of $\stmt_2$ given $\stmt_1$ is
@@
 Suppose that, given a process, we pick $t_1, t_2 \in T$.
-Given an initial description $\stmt_1 \in \tdomain_{t_1}$ and a final description $\stmt_2 \in \tdomain_{t_2}$, the \textbf{forward transition probability} $P(\stmt_2 | \stmt_1)$ tell us what is the fraction of evolutions starting in $\stmt_1$ (i.e.
-are compatible to $\stmt_1$)) that end in $\stmt_2$.
-In the same way, we can look at the \textbf{backwards transition probability}.
-The probability space is enough to recover the process, which means \textbf{forward and backward transition probabilities can always be used to fully characterize a process.}
+Given an initial description $\stmt_1 \in \tdomain_{t_1}$ and a final description $\stmt_2 \in \tdomain_{t_2}$, the \textbf{forward transition probability} $P(\stmt_2 | \stmt_1)$ tells us the fraction of evolutions starting in $\stmt_1$ (i.e.
+that are compatible with $\stmt_1$) that end in $\stmt_2$.
+In the same way, we can look at the \textbf{backward transition probability}.
+The transition probabilities are fully determined, through the measure, by the fineness relationship.
+Conversely, the fineness relationship can be, through the measure, recovered by the probabilities.
+Therefore \textbf{forward and backward transition probabilities are enough to fully characterize the behavior of a system within a process.}
 
-We say that a process is \textbf{isoentropic} if for any possible trajectory $x(t)$ we have $S_{\stmt[u]}(x(t)) = S_{\stmt[u]}(x(t + \Delta t)) \neq 0$.
-We saw that all deterministic and reversible processes are isoentropic, but the converse is not true.
+We say that a process is \textbf{isentropic} if for any possible trajectory $x(t)$ we have $S_{\stmt[u]}(x(t)) = S_{\stmt[u]}(x(t + \Delta t)) \neq 0$.
+We saw that all deterministic and reversible processes are isentropic, but the converse is not true.
 We can give, though, a characterization in terms of transition probabilities.
@@
 	\mu_{\stmt[u]}(x(t) &\AND x(t+\Delta t)) \\
-	&=  \mu_{\stmt[u]}(x(t)) P(x(t)|x(t+\Delta t)) \\
-	&=  \mu_{\stmt[u]}(x(t+\Delta t)) P(x(t+\Delta t)|x(t)) \\
+	&=  \mu_{\stmt[u]}(x(t+\Delta t)) P(x(t)|x(t+\Delta t)) \\
+	&=  \mu_{\stmt[u]}(x(t)) P(x(t+\Delta t)|x(t)) \\
 \end{align*}
@@
 \begin{equation}
-\frac{\mu_{\stmt[u]}(x(t))}{\mu_{\stmt[u]}(x(t+\Delta t))} = \frac{P(x(t+\Delta t)|x(t))}{P(x(t)|x(t+\Delta t))}.
+\frac{\mu_{\stmt[u]}(x(t+\Delta t))}{\mu_{\stmt[u]}(x(t))} = \frac{P(x(t+\Delta t)|x(t))}{P(x(t)|x(t+\Delta t))}.
 \end{equation}
 If the forward and backward probabilities are the same, the initial and final snapshots have the same measure.
-Conversely, if the initial and final have the same measure, the forward and backward probabilities must be the same.
+Conversely, if the initial and final snapshot have the same measure, the forward and backward probabilities must be the same.
 Equal measure corresponds to equal process entropy.
-Therefore \textbf{isoentropic processes are those and only those for which backward and forward probabilities are the same}.
+Therefore \textbf{isentropic processes are those and only those for which backward and forward probabilities are the same}.
 
@@
 		\hline 
-		$S_{\stmt[u]}(\stmt)$ & Evolution entropy & Defined as $\log \mu_{\stmt[u]}(\stmt)$ represents the number of bits that separate the level of description of $\stmt$ from the one provided by the unit statement $\stmt[u]$; it quantifies the number of questions needed to go from the level of description of $\stmt$ to the level of description of $\stmt[u]$  \\
+		$S_{\stmt[u]}(\stmt)$ & Evolution entropy & Defined as $\log \mu_{\stmt[u]}(\stmt)$; represents the number of bits that separate the level of description of $\stmt$ from the one provided by the unit statement $\stmt[u]$; it quantifies the number of questions needed to go from the level of description of $\stmt$ to the level of description of $\stmt[u]$  \\
 		\hline 
-		$P(\stmt_2 | \stmt_1)$ & Probability & Defined as $\mu_{\stmt_1}(\stmt_1 \AND \stmt_2)$ quantifies the ratio of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$  \\ 
+		$P(\stmt_2 | \stmt_1)$ & Probability & Defined as $\mu_{\stmt_1}(\stmt_1 \AND \stmt_2)$; quantifies the fraction of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$  \\ 
 		\hline 
@@
 		\hline 
-		 & Finitely comparable & Two statements are finitely if the measure of one in terms of the other is finite and non-zero \\ 
+		 & Finitely comparable & Two statements are finitely comparable if the measure of one in terms of the other is finite and non-zero \\ 
 		\hline 
@@
 
-\subsection{Thermodynamic processes}
+\section{Thermodynamic processes}
 

### 2020-09-21 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Initial skeleton for phil paper with Josh

\documentclass[letterpaper]{article}

\begin{document}

\title{TBD}
\author{TBD}

\date{\today}

\maketitle

\begin{abstract}
	TBD
\end{abstract}

\section{Introduction}

TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.

\subsection{Define Lagrangian}

Talk about disippative systems.

\subsection{Fundamentality}

Different notions of fundamentality.

Joint carving, perfectly natural properties. Ontological notion of fundamental. Clearly state why this is not "good" for this context. We are agnostic. One may use following points to argue whatever they want.

Joint carving? What notion of fundamentality? Hamiltonian mechanics is "fundamental" because allows to show that densities are coordinate independent quantities and therefore physical quantities. Canonical pairs (position and conjugate momentum) allows a simpler expression/comparison of densities over space. That is why statistical mechanics works best on phase space (we have a uniform measure over q/p).

Hamiltonian mechanics makes fewer assumptions on the physical system being studied than Lagrangian mechanics. Lagrangian is a subset than Hamiltonian. Makes "amount of structure" more rigorous. Hamiltonian defines the relationship between conjugate momentum and velocity ($dq/dt = \partial H / \partial p$), so kinematic assumption constrains the Hamiltonian.

The action principle can be understood geometrically in the extended phase space as a consequence of "state density" conservation. It also holds when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum. Also shows that Lagrangian is "unphysical" and depends on gauge.

\bibliography{bibliography}

\end{document}

### 2020-09-28 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Just testing to make sure that I can edit things and do pull requests
Update classical-phil-article.tex
Testing second time

@@
 
+%% testing testing! 
+%% testing twice
+
 TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.
 
-\subsection{Define Lagrangian}
+\subsection{Define Lagrangian \& Hamiltonian mechanics}
 
@@
 
-Different notions of fundamentality.
+Different notions of fundamentality, perfectly natural properties, naturalness, etc.
 
@@
 
-Joint carving?
-What notion of fundamentality?
-Hamiltonian mechanics is "fundamental" because allows to show that densities are coordinate independent quantities and therefore physical quantities.
-Canonical pairs (position and conjugate momentum) allows a simpler expression/comparison of densities over space.
-That is why statistical mechanics works best on phase space (we have a uniform measure over q/p).
+Joint carving? What notion of fundamentality?
+
+\section{Argument for privileging Hamiltonian mechanics}
+
+\subsection{Relevant formal background from assumptions of physics framework}
 
@@
 
-The action principle can be understood geometrically in the extended phase space as a consequence of "state density" conservation.
+Hamiltonian mechanics is "fundamental" because allows to show that densities are coordinate independent quantities and therefore physical quantities.
+Canonical pairs (position and conjugate momentum) allows a simpler expression/comparison of densities over space.
+That is why statistical mechanics works best on phase space (we have a uniform measure over q/p).
+
+Can characterize this argument as physically or epistemically privileging Hamiltonian mechanics
+
+\section{Interpreting Lagrangian mechanics}
+
+Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics.
+Curiel frequently appeals to naturalness and a distinction between natural vs.
+ad hoc/artificial constructions [figure out what his definition of ``naturalness" is, if he provides one].
+
+For instance, Curiel argues that Hamiltonian mechanics is less natural and thus less physically significant/privileged:  ``Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
+Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2)" (306).
+
+Curiel's main point: the representation of a classical mechanical system in the Lagrangian framework is \textit{natural} in a way that its representation in a Hamiltonian formulation is not.
+The former representation depends on intrinsic structures of the classical mechanical system rather than an artificial construction (270).
+
+Part of our response: The action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
 It also holds when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
-Also shows that Lagrangian is "unphysical" and depends on gauge.
+Also shows that Lagrangian is ``unphysical" and depends on gauge.
+
+Upshot: can better understand Lagrangian mechanics as a special case of Hamiltonian mechanics.
+Explain how the geometrical understanding of the stationary action in extended phase space should meet any criterion of ``naturalness" or ``natural construction." Lagrangian mechanics is actually less natural because contingent on having a bijection between velocity and conjugate momentum.
 

### 2020-10-05 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Added notes for response to Curiel
Adding note on strong/weak assumption

@@
 
+\usepackage{amsmath}
+
 \begin{document}
+
+%% testing from desktop
 
@@
 Talk about disippative systems.
+
+\subsubsection{Strong vs weak Kinematic assumption}
+
+Problem: changing variables from state variables $(q^i, p_j)$ to kinematic variables $(x^i, v^j)$.
+We set $x^i=q^i$.
+The change of name helps us remember that $\frac{\partial}{\partial x^i} \neq \frac{\partial}{\partial q^i}$ as one is taken at constant $v$ while the other is taken at constant $p$.
+
+Weak assumptions: $v^i = v^i(q^j, p^k)$ invertible. Strong assumptions: $m v^i = g^{ij} (p_j - q A_j(q^k))$ linear in $p_j$.
+
+We want to transport the density from state variables to initial conditions: $\rho(q^i, p_j) = |J| \rho(x^i, v^j)$. We have:
+\begin{equation}
+	|J| = \begin{bmatrix}
+	\frac{\partial x^i}{\partial q^j} & \frac{\partial v^i}{\partial q^j} \\
+	\frac{\partial x^i}{\partial p_j} & \frac{\partial v^i}{\partial p_j}
+	\end{bmatrix}
+	= \begin{bmatrix}
+	\delta^i_j & \frac{\partial v^i}{\partial q^j} \\
+	0 & \frac{\partial v^i}{\partial p_j}
+	\end{bmatrix}
+	= \left|\frac{\partial v^i}{\partial p_j}\right|
+\end{equation}
+Under the strong assumption, $\left|\frac{\partial v^i}{\partial p_j}\right| = \frac{1}{m} |g^{ij}|$.
 
@@
 
+Response: symplectic geometry indeed does not care about the difference between q's and p's.
+Classical mechanics does.
+Misindentification of the math (symplectic geometry) with the physics (Classical Hamiltonian mechanics).
+
 Curiel's main point: the representation of a classical mechanical system in the Lagrangian framework is \textit{natural} in a way that its representation in a Hamiltonian formulation is not.
@@
 
+%
+
 \bibliography{bibliography}

### 2020-10-19 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Adding more notes
Adding diagram and synopsis of derivation

--- 2020-09-28 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 
+\usepackage{amsmath}
+\usepackage{amsfonts}
+\usepackage{graphicx}
+
 \begin{document}
+
+%% testing from desktop
 
@@
 Talk about disippative systems.
+
+\subsection{Derivation}
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{Diagram.png}
+\end{figure}
+
+\subsubsection{Infinitesimal reducibility}
+
+Classical phase space is recovered under the:
+Infinitesimal reducibility assumption: the state of the system is reducible to the state of its infinitesimal parts.
+That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the state of its parts and so on.
+
+Let $\mathcal{C}$ be the state space for the whole system.
+We call particle the limit of recursive subdivision.
+Let $\mathcal{S}$ be the state space for the particles.
+Then for each $c \in \mathcal{C}$ we can find one and only one $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of its parts.
+That is, the state of the whole is a distribution over the infinitesimal parts.
+If $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space.
+
+The critical issue is that, on a manifold, $\rho_c$ must transform both as a scalar function (the value must depend on the point and not on the coordinates) and a density.
+Classical phase space (i.e.
+a symplectic manifold) is the only space that have this structure.
+Let's see why.
+
+We call state variables a set of quantities $\xi^a$ (i.e.
+coordinates in manifold terminology) that fully identify a state.
+We call coordinate $q \in \xi^a$ a particular variable that defines a unit.
+Suppose the particle state space is such that one coordinate is sufficient to define the unit system.
+That is, $\xi^a = \{ q, k^i \}$.
+We need to make sure that we can arbitrarily change the unit without changing $\rho_c$.
+We also have to make sure that $q$ fully defines the units for all state variables.
+How many $k^i$ can we have?
+
+Suppose we change unit $\hat{q}=\hat{q}(q)$.
+Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
+We have $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$.
+Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must be equal to 1.
+Note that, since $\hat{q}$ depends only on $q$, $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$.
+Suppose there is only one variable.
+Then we would have $\left|\frac{\partial \hat q}{\partial q} \right| = 1$.
+But this would mean the unit change cannot be arbitrary.
+Therefore we must have two or more variables and therefore $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$.
+This puts a constraint only on the determinant of the transformation.
+Suppose there are three or more variable.
+This constraint is not enough to recover the transformation uniquely and therefore $\hat q$  would not fully define the units for all other state variable.
+This means there must be two variable: $q$ and a single $k$.
+Coordinate independent areas and densities can only be defined on a two-dimensional manifold.
+
+To generalize, we say two coordinates are independent if changing the units for one does not change the units for the other.
+Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$.
+Suppose you change the first coordinate $q^1$ without changing the others.
+Then we will find a variable $k_1$ that changes as before so that the densities are invariant.
+Now change the second coordinate $q^2$ in the same way while also fixing $k_1$.
+Then we find a correspond $k_2$.
+In the end, we will find that $\mathcal{S}$ is $2n$-dimensional and the state variables are $\xi^a = \{ q^i, k_i \}$.
+We define an independent degree of freedom the space charted by a pair of such variables.
+
+We can now show that $\mathcal{S}$ is a symplectic manifold.
+Invariant densities can only be assigned to infinitesimal areas.
+Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define the integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
+Because degrees of freedom are, at least locally, independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
+This means the volume form is given by $\omega^n$.
+This cannot be degenerate, or some volumes would not contain states.
+Therefore $\omega$ itself cannot be degenerate.
+As the degrees of freedom are independent, the number of states on a surface does not change if we translate it across independent degrees of freedom.
+If we imagine a parallelopiped, the integral over the surface must be zero (integral over opposite sides are equal and opposite).
+Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate and therefore is a symplectic manifold.
+By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
+
+The conclusion is that, under the infinitesimal reducibility assumption, the state space of the system is a coordinate invariant distribution over the state space of the infinitesimal parts.
+If the state space of the particles is a manifold, then it is a symplectic manifold, which is the only type of manifold that supports coordinate invariant distributions.
+
+\subsubsection{Deterministic and reversible evolution}
+
+Deterministic and reversible evolution assumption: given the present state of the system, all future (determinism) and past (reversibility) states are uniquely identified.
+
+In our case, this goes to both the state of the whole and to the state of the parts.
+That is, given an initial state $s_0$ there will be one and only one evolution $s(t)$ such that $s(t_0) = t_0$.
+Moreover, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$ the density remains the same throughout the evolution.
+That is, all particles that start with the initial state must end up in the same final state and vice-versa.
+
+Now consider the integration $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
+The region $\Sigma$ will be mapped to a new region, but the fraction of the system found in both will be the same.
+If both the integral and $\rho_c$ have to remain during the evolution, then $\omega$ will also need to remain the same.
+That is, the areas in phase space must be mapped to areas of equal size and independent degrees of freedom must be mapped to independent degrees of freedom (or volumes would not be mapped to equal volumes).
+The evolution is a symplectomorphism and corresponds to Hamiltonian evolution.
+Intuitively, this is the inverse of Liousville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
+
+\subsubsection{Motion equivalence}
+
+Having recovered classical Hamiltonian mechanics, we can further restrict the law of evolution with the following:
+
+Motion equivalence assumption: the motion of the system (i.e. trajectories in physical space-time) is enough to recover its dynamics (i.e. evolutions in state space) and vice-versa.
+
+If we focus on the parts, this means that for every evolution in phase space there should be one and only one trajectory.
+Note that each space variable $x^i$ is a coordinate, a state variable that defines a unit.
+In fact, the trajectories can be fully described by those units and only those units.
+So we can say that $q^i=x^i$, each $q^i$ will be paired with a conjugate $p_i$ and each state $\{q^i, p_i\}$ will be mapped to one and only one trajectory.
+At each point $x^i$, then, must pass infinitely many trajectories, one for each combination of $\{p_i\}$.
+Since the trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
+If the equations of motions were such that $v^i=v^i(q^i)$, then kinematic equivalence would fail as the full trajectory would be determined only by $q^i$.
+So we must have $v^i=v^i(q^i, p_i)$.
+The relationship must be invertible or kinematic equivalence would fail.
+At any given time, then, we must the following relationship:
+\begin{equation}
+\begin{aligned}
+x^i &= q^i \\
+v^i &= d_t x^i = v^i(q^j, p_k)
+\end{aligned}
+\end{equation}
+
+In this case, that we call weak equivalence, $v^i(q^j, p_k)$ must be invertible at every $q^i$.
+Therefore Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
+Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ must be non-zero everywhere, and therefore must have the same sign which we take to be positive.
+This are exactly the cases where a Lagrangian can be constructed, and are exactly the cases where the Lagrangian leads to a unique solution.
+
+If we look at the whole distribution, we have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
+\begin{equation}
+|J| = \begin{bmatrix}
+\frac{\partial x^i}{\partial q^j} & \frac{\partial v^i}{\partial q^j} \\
+\frac{\partial x^i}{\partial p_j} & \frac{\partial v^i}{\partial p_j}
+\end{bmatrix}
+= \begin{bmatrix}
+\delta^i_j & \frac{\partial v^i}{\partial q^j} \\
+0 & \frac{\partial v^i}{\partial p_j}
+\end{bmatrix}
+= \left|\frac{\partial v^i}{\partial p_j}\right|.
+\end{equation}
+Note that while the value given by $\rho(q^i, p_j)$ is coordinate independent, the value given by $\rho(x^i, v^j)$ depends on the choice of coordinate through $\left|\frac{\partial v^i}{\partial p_j}\right|$.
+If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only.
+Similar considerations will also hold for marginal distributions (i.e.
+distribution on a subset of the coordinates) which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
+We set:
+\begin{equation}
+\begin{aligned}
+\frac{\partial v^i}{\partial p_j} = \frac{1}{m} g^{ij} \\
+\frac{\partial p_j}{\partial v^i} = m g_{ji}
+\end{aligned}
+\end{equation}
+where $m$ is the unit conversion constant between velocity and conjugate momentum while $g_ij$ represents the linear dependency.
+
+If we integrate, we have:
+\begin{equation}
+\begin{aligned}
+v^i = \frac{1}{m} g^{ij}(p_j - A_j) \\
+p_j = m g_{ji} v^i + A_j
+\end{aligned}
+\end{equation}
+where $A_j$ are arbitrary functions. Note that:
+\begin{equation}
+v^i = d_t q^i = \frac{\partial H }{\partial P} = \frac{1}{m} g^{ij}(p_j - A_j) \\
+\end{equation}
+We integrate yet again and find:
+\begin{equation}
+H = \frac{1}{2m} (p_j - A_j) g^{ij}(p_j - A_j) + V \\
+\end{equation}
+where V is another arbitrary function. We recognize this as the Hamiltonian for massive particles under potential forces.
+
+Therefore
 

### 2020-10-23 Articles/2019-Entropy/Source/entropy-article.tex
More work on the core ideas

@@
 
-\section{Evolutions, states and levels of description}
+\section{Introduction}
+
+\emph{NB: this is a draft.}
+
+The basic premise of statistical mechanics is that entropy counts states.
+We believe this premise is not correct and what entropy really counts evolutions (i.e.
+curves in phase space parameterized by time).
+If the evolution is deterministic and reversible (i.e.
+each state corresponds to one evolution), these notions coincides, but leads to confusion in the rest of the cases.
+
+Consider this simple example: the system at time $t_0$ can be in state $a$ or $b$.
+In both cases, however, the system will move to state $a$ at time $t_1$.
+The dynamics is deterministic: given the state at $t_0$ we can tell the state at time $t_1$.
+Yet, the dynamics is non-reversible: the states have merged into the equilibrium $a$, we can't reconstruct the previous state, we have lost information.
+
+Note that the count of states goes down over time.
+If we start at time $t_0$ with the set with both $a$ and $b$, we have to states.
+Through the map, this becomes the set with only $a$ at time $t_1$.
+We go from two to one.
+If we say entropy count states, then we are saying entropy decreases.
+
+On the other hand, we propose that entropy counts the evolutions per state.
+In this case, the initial set of two states and the final set of one state corresponds to the same number of possible evolutions, two.
+However, the number of evolutions corresponding to each state changes: at time $t_0$ we have one evolution per state; at time $t_1$ we have two evolutions per state.
+Our entropy has increased.
+
+Note that, whenever evolutions merge, the dynamics is irreversible, the entropy increases.
+In the continuous case, evolutions don't need to merge, just become more dense.
+At equilibrium, evolutions will no longer merge: the entropy will become stable.
+This gives a very simple explanation of why entropy increases.
+The rest of the paper aims to make this simple intuition more precise.
+
+To achieve our goal, there are a number of issues that needs to be solved.
+First of all, we need to understand exactly how states are defined and how are they assigned an entropy.
+In fact, note that in the example the state $a$ has different entropy at different times.
+While at first this may seem strange, it actually makes sense: saying that the system is in state $a$ at time $t_0$ is different than saying that it is in state $a$ at time $t_1$: in the first case, we know the system couldn't be in state $b$ at time $t_0$.
+Given the non-reversible dynamics, the same assertion at different times may provide different information.
+
+Our starting point, then, needs to be a framework that can keep track of different descriptions at different times with different level of granularity.
+Within this general framework, we need to understand what are the conditions that allow us to group some into descriptions of the same system.
+This, it turns out, is not so trivial and comes with some necessary assumptions.
+
+Lastly, we want to understand what additional assumptions are needed to recover the different physical frameworks, such as classical mechanics, statistical mechanics and, our main focus, thermodynamics.
+The hope is that building everything from the ground up will give us more insight in the foundations of all the respective theories and, more in general, of physics.
+
+Note that our aim here is not to present a fully fleshed our mathematical framework. We want simply to present a blueprint for it and convince the reader that this is doable and interesting.
+
+\section{A general framework for physical processes}
 
@@
 This can be thought of as the number of bits, yes/no questions, that separate the level of description of $\stmt[u]$ from the one provided by $\stmt$.
-Connecting the definitions from above, under a deterministic process we have $S_{\stmt[u]}(x(t)) \leq S_{\stmt[u]}(x(t + \Delta t))$ since $x(t)\narrower x(t+\Delta t)$ means $x(t) \finer x(t+\Delta t)$ and $\mu_{\stmt[u]}(x(t)) \leq \mu_{\stmt[u]}(x(t+\Delta t))$.
+Connecting the definitions from above, under a deterministic process we have
+\begin{equation}\label{entropy_increases}
+S_{\stmt[u]}(x(t)) \leq S_{\stmt[u]}(x(t + \Delta t))
+\end{equation}
+since $x(t)\narrower x(t+\Delta t)$ means $x(t) \finer x(t+\Delta t)$ and $\mu_{\stmt[u]}(x(t)) \leq \mu_{\stmt[u]}(x(t+\Delta t))$.
 If the process is determistic and reversible, we have $S_{\stmt[u]}(x(t)) = S_{\stmt[u]}(x(t + \Delta t))$.
@@
 
+At this point it would seem the state entropy has a relative scale like the process entropy.
+However, system composition makes the scale absolute.
+To see this, let us define a \textbf{null system} as one with only one possible ``state'' $x_0$, which corresponds to the fact that there is no system at all.
+We call this the \textbf{null state}.
+The null system functions as a zero for system composition:  any system composed with the null system gives the original system.
+We are therefore forced to use $x_\emptyset$ as a unit (if not because it is the only statement with non-zero measure) and have:
+\begin{equation}\label{entropy_zero}
+S_{x_\emptyset}(x_\emptyset) = \log \mu_{x_\emptyset}(x_\emptyset) = \log 1 = 0 \\
+\end{equation} 
+Conceptually, if the null system is composed with an independent system, the total possible evolutions are the ones coming from the independent system.
+
+Since we said before that a system may not be present in all processes, to make the state space both conceptually and mathematically complete, it will have to contain a null state.
+The state entropy of the null state of each system must be zero, for exactly the same reasons as before.
+The state entropy, then, is always taken with respect to the null state, for which it is zero and will be noted simply by $S$.
+Also note that, all else being equal, saying that a system is in a particular state cannot give us a more refined description than saying that a system is missing.
+While there may be more than one way that a particular state is implemented (e.g.
+degenerate energy level, multiple microstates leading to the same macrostate, ...), there is only one way the system can be missing.
+If we know the system is missing, there is nothing else that could be said about it.
+Therefore the description provided by the null state is always finer than any other description of the system, including any state.
+For any state $x \in X$ we have:
+\begin{equation}\label{entropy_lower_bound}
+\begin{aligned}
+x_\emptyset &\finer x \\
+S(x) &\geq S(x_\emptyset) = 0
+\end{aligned}
+\end{equation}
+In other words, \textbf{the null system and the null states set an absolute scale for the state entropy}.\footnote{The fact that each state space includes a null state, and that the entropy of the null state is zero, is required by the math to make $\iota$ work in the case where the system is not present.
+In fact, in that case all the states are impossible none of them would map to the snapshot of the (missing) system.
+As of now, it is not yet clear whether the fact that the entropy of any state must be equal or greater to the one of the null state is an additional mathematical requirement or not.}
+
 We say $\xi^a : X \to \mathbb{R}$ are a set of state variables if they uniquely identify a state.
@@
 We'll see that, for thermodynamic systems, this will correspond to the usual thermodynamic entropy.
-We define a \textbf{purely mechanical} system one for which the state entropy is the same across all states (i.e.
-$S=k$ where $k$ is a constant, typically chosen to be zero).
+We define a \textbf{purely mechanical} system one for which the state entropy is zero across all states.
 We'll see that classical phase space and quantum state spaces are state spaces of purely mechanical systems.
@@
 
-\section{Thermodynamic processes}
+\section{Mechanics}
+The recovery of classical mechanics is something that we have already covered in previous works.
+In fact, this goal of this work is to try and extend the framework to be applicable to non-derministic and non-reversible processes.
+
+The fundamental assumption to recover classical mechanics is that \textbf{the processes describes the deterministic and reversible evolution of a single system composed by infinitesimal parts, each described by a set independent degrees of freedom.} In this case, each evolution corresponds to one and only one state at each time.
+The state space and each time domain, then, can be given a uniform measure.
+The system is purely mechanical.
+Each degree of freedom is described by a coordinate $q^i$ and a conjugate variable $k_i$ of inverse units, making the areas $dq^idk_i$ invariant.
+Hamiltonian mechanics is recovered by requiring that the evolution is deterministic and reversible and the independence of the degrees of freedom is preserved.
+This dynamics is the one and the only one that preserves state entropy.
+
+What is often missed is that the assumption of determinism and reversibility is crucial to be able to give the same measure over all states at all times.
+Suppose we have a dynamical system that exhibits equilibria (i.e.
+attractors).
+Each equilibria will have a basin of attraction.
+We can assume, like in the previous cases, that at a particular time all states are possible and each state corresponds to one evolution.
+That is, we have a uniform measure and a uniform entropy on that particular time domain.
+If we let time pass, however, the possible states will shrink around the equilibria.
+
+\section{Thermodynamics}
 
@@
 \end{equation}
-which recovers the second law.
+which recovers the second law of thermodynamics. This is just a more specialized version of \eqref{entropy_increases} which is valid for any deterministic process.
+
+Recall that, by \eqref{entropy_lower_bound}, the state entropy $S(U)$ must be bounded from below.
+If the temperature tends to zero for some state $U_0$, the curve is tangent to a vertical line at $U_0$, so the entropy will be bound to a positive finite value.
+This recovers the third law.
+
+While we technically recovered the letter of the law, we do not believe have really recovered the spirit.
+In fact, the system may have a lower the curve could continue, have lower values of $U_0$ with even lower values of $S$.
+We could also have states for negative values of $U$.
+The law does not say anything on the subject.
+There are, in fact, other facts that are usually implied but not stated: that the temperature will decrease when energy decreases, that there will be a minimum value for energy and that, for that value, temperature will tend to zero.
+
+If we traverse $S(U)$ from higher
+
+In \eqref{entropy_zero} we saw that the state entropy must be zero or greater. It is zero if the state corresponds to one independent evThe third law comes in many forms, so
 

### 2020-10-26 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Incorporating some draft paragraphs on content related to North 2009 and Curiel 2014.

@@
 
-\usepackage{amsmath}
+\usepackage{amsmath, hyperref}
 \usepackage{amsfonts}
 \usepackage{graphicx}
+
+\usepackage[backend=biber,  style=chicago-authordate, url=false, doi=false, eprint=false]{biblatex}
+%\usepackage[backend=biber,  style=authoryear-icomp, url=false, doi=false, eprint=false]{biblatex}
+\addbibresource{bibliography.bib}
+
+\AtEveryBibitem{%
+  \clearfield{note}%
+  \clearfield{series}%
+  \clearfield{pages}%
+}
 
@@
 
-%% testing testing! 
-%% testing twice
-
 TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.
@@
 Talk about disippative systems.
+
+Throughout, we will focus only on non-dissipative classical systems, i.e.
+conservative systems.
+For in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $.
+This requires that the system is conservative [why exactly?], which is equivalent to the system being deterministic and reversible.
+Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
+% Dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential
 
@@
 \end{figure}
+
+Overall summary: A classical system is an infinitesimally reducible system.
+The state space of these systems can be modeled by a symplectic manifold (the physical assumption of infinitesimal reducibility corresponds to the mathematical structure of symplectic manifolds).
+Classical Hamiltonian systems satisfy the additional conditions of having deterministic and reversible evolutions.
+These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it contact manifold.
+Classical Lagrangian systems satisfy a further assumption: keeping track of the trajectories in physical space is equivalent to tracking the evolution of states in phase space.
+Note that it is not always true that we can determine the evolution in phase space from the trajectory in physical space.
+For instance, knowledge of the trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from its trajectory alone.
 
@@
 
-Therefore
-
-\subsection{Fundamentality}
-
-Different notions of fundamentality, perfectly natural properties, naturalness, etc.
-
-Joint carving, perfectly natural properties.
-Ontological notion of fundamental.
-Clearly state why this is not "good" for this context.
-We are agnostic.
-One may use following points to argue whatever they want.
-
-Joint carving? What notion of fundamentality?
-
 \section{Argument for privileging Hamiltonian mechanics}
@@
 
+%On the notion of more or less structure, treated in such a way that we avoid the objections of Halvorson/Swanson against North's notion of amount of structure:
+Our comparisons of structure arise from logical chains of strict entailment, wherein something has more structure if and only if it is less general, i.e.
+if and only if it is a specialization of a more general case.
+So in this case, the Lagrangian framework has more structure than the Hamiltonian framework because Lagrangian mechanics requires an additional assumption.
+This difference in the number of assumptions corresponds to a difference in the number of commitments.
+When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?).
+In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.
+
 Hamiltonian mechanics makes fewer assumptions on the physical system being studied than Lagrangian mechanics.
 Lagrangian is a subset than Hamiltonian.
-Makes "amount of structure" more rigorous.
+Makes ``amount of structure" more rigorous.
 Hamiltonian defines the relationship between conjugate momentum and velocity ($dq/dt = \partial H / \partial p$), so kinematic assumption constrains the Hamiltonian.
 
-Hamiltonian mechanics is "fundamental" because allows to show that densities are coordinate independent quantities and therefore physical quantities.
+Claim to think about: GC was arguing on 10-12 that you could handle classical systems in the Hamiltonian framework with just a differential manifold and no symplectic structure.
+This would contradict North's indispensability claim that Hamiltonian symplectic structure is necessary for treating classical systems.
+E.g., North footnote 31: ``The symplectic form is crucial to determining the motion, since the same Hamiltonian can induce different flows for different symplectic forms" [76].
+North takes this as an argument for why we cannot eliminate the symplectic form in an attempt to get at even less structure.
+We need it to define the dynamics.
+
+\subsection{The significance of invariant distributions}
+
+North argues that Hamiltonian mechanics is superior because it requires strictly less structure than Lagrangian mechanics.
+While we agree that Hamiltonian mechanics does require less structure---in a different sense than North's account---this is not the only grounds for privileging it.
+Instead, the main grounds for privileging Hamiltonian mechanics comes from understanding physics in terms of invariant distributions.
+This allows us to provide a better physical interpretation of the action principle, which holds even when a Lagrangian cannot be defined for the system of interest.
+% Do we agree with the following claim defended by North: ``as far as we can tell, we need only symplectic structure to do classical mechanics" [74].
+
+Hamiltonian mechanics is ``fundamental" because allows to show that densities are coordinate independent quantities and therefore physical quantities.
 Canonical pairs (position and conjugate momentum) allows a simpler expression/comparison of densities over space.
@@
 
-\section{Interpreting Lagrangian mechanics}
+\subsection{Response to Curiel's criticisms of Hamiltonian mechanics}
 
-Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics.
-Curiel frequently appeals to naturalness and a distinction between natural vs.
-ad hoc/artificial constructions [figure out what his definition of ``naturalness" is, if he provides one].
+Curiel restricts Hamiltonian mechanics to systems whose Hamiltonian takes the form (equation 6.2) $H =\alpha^{m n} p_m p_n + U (q_j) $.
+Here, the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q $.
+Curiel's argument in a nutshell:  ``Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
+Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2)" (306).
 
-For instance, Curiel argues that Hamiltonian mechanics is less natural and thus less physically significant/privileged:  ``Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
-Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2)" (306).
+Curiel argues that Hamiltonian mechanics is less physically significant than Lagrangian mechanics.
+His argument hinges on the claim that Hamiltonian mechanics requires an ad hoc condition that is imposed by hand, whereas Lagrangian mechanics follows from physically plausible principles.
+We agree with Curiel that his equation 6.2 requires a condition that is imposed by hand, namely the kinematic equivalence assumption.
+However, this condition has a clear physical meaning, and it corresponds to a physical condition that must hold in order to define/apply Lagrangian mechanics.
+Pace Curiel, there exist Hamiltonian systems that do not satisfy his equation 6.2 (equivalently, classical systems where the momentum is not a linear function of the velocity), and these will even have a Lagrangian, although they will not be an abstract classical system in Curiel's sense.
+Hence, there seems to be a burden on Curiel to distinguish---in a principled way---the classical Lagrangian systems from those that are not classical.
+% Hence, perhaps we agree with Curiel that ``it is not necessary that [these conditions] hold in order for the theory to be meaningfully applied to model a [classical] system" [305].
+Where is Curiel takes this as an argument for the primacy of Lagrangian mechanics, we take it as an argument for the primacy of Hamiltonian mechanics, since Curiel's conditions are equivalent to imposing motion equivalence.
 
@@
 
+On its own, the kinematic/motion of a system in physical space is not sufficient for defining the system's dynamics.
+The kinematics/motion does not suffice to determine the forces acting on the system.
+For it is essential to be able to distinguish the forces acting on an inertial system from fictitious forces arising in the description of a non-inertial system.
+Hence, unless one knows that the kinematic equivalence principle is satisfied, Lagrangian mechanics faces an underdetermination problem: it is unable to distinguish between certain inertial vs.
+non-inertial systems that are physically distinct.
+Distinguishing between these kinds of systems requires a relationship between conjugate momentum and kinematic momentum (or equivalently, knowledge of the energy/Hamiltonian).
+% Example of underdetermination in the context of Lagrangian mechanics: a free particle in a non-inertial frame vs.
+an accelerated particle (in an inertial frame) acted on by a dissipative force.
+These two distinct classical systems can nevertheless have the same equations of motion.
+% The Hamiltonian framework does not face a corresponding underdetermination problem because it does not allow for dissipative systems.
+Hence, there is a sense in which the Lagrangian formalism is more mathematically general than the Hamiltonian formalism.
+However, this additional generality is irrelevant when considering conservative classical systems.
+
+The standard predator--prey model from differential equations satisfies Curiel's definition of an abstract classical system.
+However, it does not satisfy Curiel's notion of the split between configurational vs.
+velocital coordinates.
+And it is not true that the velocity will be the first derivative of the position.
+Moreover, the predator--prey system is neither Hamiltonian nor Lagrangian, because it is not necessarily deterministic and reversible.
+% Related to equation 3.1. Make precise.
+% Furthermore, the momentum does not need to be a linear function of velocity. For example, the photon does not satisfy this relationship.
+
+Objection against Curiel to make precise and substantiate with textual evidence: implicitly, Curiel assumes that classical systems are defined as Newtonian systems in an inertial frame with Cartesian coordinates.
+Implicitly, Curiel assumes that energy is a function of position and velocity, entailing that we have a Hamiltonian.
+The claim in footnote 17 that the kinetic energy is a function of velocity depends on the choice of a particular coordinate system; it is not a coordinate-invariant claim.
+Curiel's thought experiment assumes that we have a constant force (in this case gravity) and that we are in a reference frame where we can compare forces (e.g.
+an inertial frame).
+
+\subsection{Discussion of relation to categorical equivalence proofs}
+
+Possible objection to consider stemming from categorical equivalence: for systems in the hyper regular domain, does the categorical equivalence of Hamiltonian and Lagrangian mechanics show that we require the same number of assumptions to treat systems in this domain?
+Some relevant considerations for responding: note that the photon does not belong to the hyper regular domain.
+Furthermore, if you are not in the hyper regular domain, then there is no longer a one-to-one map between position and velocity.
+In this sense, the Hamiltonian framework is more general because it applies outside the hyper regular domain: the Hamiltonian can define a unique solution (unique evolution in phase space) even when the Hessian is zero.
+In contrast, whenever the Lagrangian is not hyper regular, the Euler--Lagrange equations do not possess a unique solution.
+Furthermore, density plays a special role in the Hamiltonian framework, whereas it is in some sense an accident that the density and configuration space is invariant (simply because the Jacobian equals unity in Cartesian coordinates).
+
+\section{A non-metaphysical notion of fundamentality}
+
+%Different notions of fundamentality, perfectly natural properties, naturalness, etc.
+
+%Joint carving, perfectly natural properties.
+Ontological notion of fundamental.
+Clearly state why this is not ``good" for this context.
+We are agnostic.
+One may use following points to argue whatever they want.
+
+Whereas North \parencites*[76]{North2009} seeks to determine the `real' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.
+Whether Hamiltonian phase space structure is metaphysically fundamental is neither here nor there for the questions we seek to answer.
+For regardless of its metaphysical status, Hamiltonian mechanics is epistemically, conceptually, and physically more fundamental.
+% Similarly, Curiel advocates neutrality on these metaphysical issues when it comes to the conceptual analysis of a physical theory's structure (304).
+% On the basis of the fundamentality of Hamiltonian mechanics, North argues that ``we should infer that symplectic structure is the `real' state space structure of the theory.
+We should infer that this structure corresponds to the `real' physical structure of a classical mechanical world" [76]
+%``there is an objective fact about what structure exists, there is a privileged carving of nature at its joints: along the lines of its fundamental physical structure" [82].
+
+Methodologically, we agree with North that it is important to identify invariants and more generally ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North2009}.
+Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North2009}.
+However, whereas North proceeds mainly from the mathematical structure to claims about physical significance, we seek to interpret the mathematics by beginning with physically significant assumptions.
+We seek to determine which physical assumptions are necessary and sufficient for supporting a corresponding mathematical structure.
+This does not entail that additional mathematical structure is necessarily physically insignificant or surplus structure, just that it may correspond to an additional physically significant assumption that might apply to the system under consideration.
+As we will see, this characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system.
+Hence, whereas North views the Lagrangian structure as surplus or redundant (``excess, superfluous structure"), we take the Lagrangian structure as unnecessary but still physically significant, corresponding to an additional physical assumption that applies to some classical systems.
+% % Similarly, we agree with the general methodology of Curiel as well: ``One task of the philosopher of physics is to work out the conceptual structure of a physical theory---or at least an appropriate one, if it is not unique---to isolate and characterize its mathematical and formal structures, and show how natural physical interpretation accrues to them from an analysis of their empirical application" (272).
+% I think that Curiel might make a similar point regarding the significance of a formulation even if it uses surplus structure or less physically significant structure.
+See section 7 argument against structural realism: Curiel argues that the Hamiltonian formulation nevertheless does not fail to represent classical systems, even if it structure is not the most natural (on his view)
+%having argued that Lagrangian mechanics has more structure ``than the theory really needs", North claims that ``we should infer that it does not exist, that it does not correspond to fundamental structure" [75].
+% Curiel on the harmonic oscillator: claims to have given an example where there is sameness of meaning (both Lagrangian and Hamiltonian formulations represent the harmonic oscillator and provide the same meaning) without sameness of structure (they're not isomorphic).
+%``we should infer that the world contains at least the amount of structure needed to formulate its fundamental dynamics in an invariant way" [80].
+% E.g., North argues that ``only certain kinds of mathematical object are appropriate for representing the structure of the world" and it is through these facts that we figure out which formulation of physics to prefer?
+[66].
+
+We agree that providing a physical interpretation of the invariants is important.
+However, some non-invariant quantities can be physically meaningful.
+For instance, the coordinates q define the units, and these are physically significant despite not being invariant.
+Units are physically significant because they allow us to assign numbers to a physical quantity.
+They are physically significant in part because they correspond to measurable quantities.
+Whereas North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North2009}, some choices of the coordinates q are less physically meaningful because they are not properly connected with measurement.
+Some arbitrary mixings of q's and p's preserve the symplectic manifold but are not as physically significant.
+For this reason, the q's are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative p's are not.
+The p's are inverses of the q's.
+Whereas symplectic geometry creates a mathematical, syntactic symmetry between the q and p variables, this is not a physically significant symmetry from the standpoint of measurement.
+Moreover: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
+(In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).
+% It follows from the way in which position quantities fix the units that $\dot{x} = v $.
+It is built into the notion of a dynamical system that half of the state variables are derivatives of the other half.
+
+In defining ``Hamiltonian mechanics" we are not privileging any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics.
+Each of these frameworks has different pragmatic costs and benefits.
+Hence, unlike North, we are comfortable acknowledging that the mathematical structure that you ``need" might not be unique.
+There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions.
+Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others.
+And in some sense, each of these three spaces involves reference to some ``unphysical" coordinates or representational means.
+Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics.
+In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.
+% One con of this extended phase space formulation: the Hamiltonian constraint is both a constraint and the generator of time evolution; hence, it must be defined in the neighborhood of the actual states, so the values it provides outside of these neighborhoods should be unphysical, yet they encode the equations of motion.
+% Should we remark on the claim that the only structure we need is whatever is sufficient to define densities locally? Hence, we need just the symplectic form but not the cotangent bundle?
+% E.g.
+in the extended phase space formulation, the equations of motion come from derivatives of the Hamiltonian constraint over the variables, but this constraint is defined on unphysical points.
+These unphysical points are plausibly a mathematical artifact, although they supply a nice tool/method for understanding physically significant aspects of the representation.
+
+\subsection{Some helpful quotations from North that could help characterize her position and relevant differences:}
+
+North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory.
+Structure comprises those quantities that remain intact when we transform one allowable coordinatization of the laws into another" [78]
+
+North definition of structure: ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" [66].
+So presumably, North would view the invariant phase space density as part of this fundamental structure.
+
+North's epistemic method for figuring out the structure: ``look at what is invariant under allowable transformations"; this tells us what the structure is [73].
+
+\section{Interpreting Lagrangian mechanics}
+
+%Discussion of gauge transformations in the context of Lagrangian vs.
+Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):
+In some sense (to be made more precise), gauge transformations are more fundamental in Lagrangian mechanics than Hamiltonian mechanics.
+The conjugate momentum $p $ functions as a dummy variable in virtue of being a gauge quantity.
+From motion equivalence, the Lagrangian has more invariant structure than the Hamiltonian [not sure what this really means].
+The Lagrangian has the advantage of being invariant under coordinate transformations, whereas the Hamiltonian is not a coordinate invariant.
+Instead, the Hamiltonian varies as the time component of a 4-co-vector.
+Nevertheless, the Lagrangian itself is physically problematic because it is a gauge-dependent quantity.
+The p variables are not of fundamental significance because we can perform a gauge transformation on the p's such that the q and $\dot{q} $ variables remain the same.
+For systems where motion equivalence holds, we are really measuring the q and $\dot{q} $ variables, indicating that this is a genuine gauge transformation of the momentum variables.
+Hence, privileging Hamiltonian mechanics does not entail, pace North, that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North2009}.
+Instead, position and kinetic momentum remain physically privileged, with conjugate momentum relegated to a gauge quantity (despite its role in the Hamiltonian formalism).
+% But there is a sense in which on the extended phase space, the Hamiltonian constraint is an invariant, yielding a coordinate-invariant way of handling systems with Hamiltonian mechanics. (?) 
+%Since the q and $\dot{q} $ variables fix the physical state, this indicates that transformed conjugate momentum is a gauge quantity. 
+%North argues that ``momentum is a truly fundamental quantity, on a par with position" [77].
+
+Claim to clarify: motion equivalence allows us to understand the gauge redundancy of the Lagrangian.
+However, it is not true that the Lagrangian has these redundancies in virtue of the system satisfying motion equivalence.
+
+Issue of the mathematical structure that is necessary/sufficient for Lagrangian mechanics:
+North footnote 20 and appendix: it is not generically true that the state space of Lagrangian mechanics comes equipped with a metric.
+Footnote 22: ``a more general, coordinate-free version" of Lagrangian mechanics ``does not make explicit use of this metric"; discussed in appendix A.
+North: ``the square of the Riemannian line element is what gives the allowable sets of coordinate transformations, the ones that preserve the Lagrangian equations of motion.
+A volume element would not suffice for this" [75].
+How does this connect with motion equivalence and its associated mathematical structure?
+
+North assumes that formulating a Lagrangian theory requires specifying a Riemannian metric; according to Curiel, this is neither necessary nor sufficient.
+Furthermore, the physical significance of a Riemannian metric in the context of a Lagrangian theory is unclear: two different Lagrangian's can have the same Lagrangian vector field and yet induce inequivalent Riemannian metrics
+
+%Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics.
+Curiel frequently appeals to naturalness and a distinction between natural vs.
+ad hoc/artificial constructions [figure out what his definition of ``naturalness" is, if he provides one].
+
 %
 
-\bibliography{bibliography}
+%\bibliography{bibliography}
+\printbibliography
 

### 2020-10-26 Overviews/Classical mechanics/ClassicalMechanicsOverview.tex
Adding classical overview

--- 2020-10-19 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
-\documentclass[letterpaper]{article}
+\documentclass[10pt,twocolumn]{revtex4-1}
+%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
+%\documentclass[10pt,twocolumn,letterpaper]{article}
 
@@
 \usepackage{graphicx}
+\usepackage{enumitem}
 
@@
 
-%% testing from desktop
-
-\title{TBD}
-\author{TBD}
+\title{Classical mechanics and infinitesimal reducibility  \\
+	(Assumptions of Physics overview)}
+\author{Gabriele Carcassi, Christine A. Aidala}
+\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
 
@@
 
+\begin{abstract}
+We briefly show how classical mechanics can be rederived and better understood as a consequence of three assumptions: infinitesimal reducibility, deterministic and reversible evolution, and kinematic equivalence.
+This work is an overview of some of the result of Assumptions of Physics (https://assumptionsofphysics.org), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+\end{abstract}
+
 \maketitle
-
-\begin{abstract}
-	TBD
-\end{abstract}
 
@@
 
-%% testing testing! 
-%% testing twice
-
-TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.
-
-\subsection{Define Lagrangian \& Hamiltonian mechanics}
-
-Talk about disippative systems.
-
-\subsection{Derivation}
+The overall argument can be summed up in the following diagram that can be used as a guide throughout this note.
 
@@
 	\centering
-	\includegraphics[width=\textwidth]{Diagram.png}
+	\includegraphics[width=\columnwidth]{Diagram.png}
 \end{figure}
 
-\subsubsection{Infinitesimal reducibility}
+The three assumptions lie on the left column.
+Each assumption lead to one or two key insights that progressively lead to the physical concepts in the middle column.
+Each of these is then mapped to its mathematical framework on the right.
 
-Classical phase space is recovered under the:
-Infinitesimal reducibility assumption: the state of the system is reducible to the state of its infinitesimal parts.
-That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the state of its parts and so on.
+\section{Infinitesimal reducibility}
 
-Let $\mathcal{C}$ be the state space for the whole system.
+\textbf{Infinitesimal reducibility assumption:} \emph{the state of the system is reducible to the state of its infinitesimal parts.} That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the state of its parts and so on.
+
+Under this assumption, the state of the whole is a distribution over the state of the parts.
+More precisely, let $\mathcal{C}$ be the state space for the whole system.
 We call particle the limit of recursive subdivision.
@@
 
-The critical issue is that, on a manifold, $\rho_c$ must transform both as a scalar function (the value must depend on the point and not on the coordinates) and a density.
-Classical phase space (i.e.
-a symplectic manifold) is the only space that have this structure.
-Let's see why.
+The next step takes the rest of the section.
+We need to show that $\mathcal{S}$ has the structure of phase space, with its conjugate variables.
+Mathematically, $\mathcal{S}$ is a symplectic manifold.
+The key insight is that, on a manifold, $\rho_c$ must transform both as a scalar function (the value must depend on the point and not on the coordinates) and as a density.
+Classical phase space is the only space that allows these invariant distributions.
 
+The key concept is to keep track of the unit system so we need precise terminology.
 We call state variables a set of quantities $\xi^a$ (i.e.
 coordinates in manifold terminology) that fully identify a state.
+That is, we can write each state $s(\xi^a)$ as a function of the state variables.
 We call coordinate $q \in \xi^a$ a particular variable that defines a unit.
-Suppose the particle state space is such that one coordinate is sufficient to define the unit system.
-That is, $\xi^a = \{ q, k^i \}$.
-We need to make sure that we can arbitrarily change the unit without changing $\rho_c$.
-We also have to make sure that $q$ fully defines the units for all state variables.
-How many $k^i$ can we have?
+The key problem is to understand how many coordinates we can have for a given set of state variables.
 
+We start with the simplest case, where one coordinates is sufficient to define the unit system, which means the following three conditions:
+\begin{enumerate}[noitemsep]
+	\item the state variables can be written as $\xi^a = \{ q, k^i \}$
+	\item we can arbitrarily change coordinate to $\hat{q}=\hat{q}(q)$
+	\item a change of coordinate induces a unique change over the remaining state variables $\hat{k}^j = \hat{k}^j(q, k^i)$.
+\end{enumerate}
+
+Now we show that there can only be one a single $k^i$.
 Suppose we change unit $\hat{q}=\hat{q}(q)$.
@@
 
-To generalize, we say two coordinates are independent if changing the units for one does not change the units for the other.
-Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$.
+Now we generalize to multiple coordinates.
+We say two coordinates are independent if changing the units for one does not induce a change the units for the other.
+Now suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$.
 Suppose you change the first coordinate $q^1$ without changing the others.
@@
 Then we find a correspond $k_2$.
-In the end, we will find that $\mathcal{S}$ is $2n$-dimensional and the state variables are $\xi^a = \{ q^i, k_i \}$.
+We can proceed in the same way until we exhaust all coordinates, which must also mean that there are no state variables left.
+We find that $\mathcal{S}$ is $2n$-dimensional and the state variables are $\xi^a = \{ q^i, k_i \}$.
 We define an independent degree of freedom the space charted by a pair of such variables.
 
-We can now show that $\mathcal{S}$ is a symplectic manifold.
-Invariant densities can only be assigned to infinitesimal areas.
-Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define the integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
+We can dress the result a bit more formally, and show that $\mathcal{S}$ is a symplectic manifold.
+To characterize marginal distributions on each degree of freedom, we want to define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$ such that the density $\rho_c$ is invariant.
+Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, and we need $\omega$ to be invariant.
 Because degrees of freedom are, at least locally, independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
@@
 
-The conclusion is that, under the infinitesimal reducibility assumption, the state space of the system is a coordinate invariant distribution over the state space of the infinitesimal parts.
-If the state space of the particles is a manifold, then it is a symplectic manifold, which is the only type of manifold that supports coordinate invariant distributions.
+Phase space (i.e. symplectic manifolds) is the only type of manifold that is able to support coordinate invariant distributions, which are required to describe an infinitesimally reducible system.
 
-\subsubsection{Deterministic and reversible evolution}
+\section{Deterministic and reversible evolution}
 
-Deterministic and reversible evolution assumption: given the present state of the system, all future (determinism) and past (reversibility) states are uniquely identified.
+\textbf{Deterministic and reversible evolution assumption:} \emph{given the present state of the system, all future (determinism) and past (reversibility) states are uniquely identified.}
 
-In our case, this goes to both the state of the whole and to the state of the parts.
-That is, given an initial state $s_0$ there will be one and only one evolution $s(t)$ such that $s(t_0) = t_0$.
-Moreover, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$ the density remains the same throughout the evolution.
-That is, all particles that start with the initial state must end up in the same final state and vice-versa.
+We first apply the assumptions to the motion of a single particle.
+That is, given a state $s_0$ at the initial time $t_0$, for each time $t$ there will be one and only one state $s_t$.
+There is one possible evolution $s(t)$ such that $s(t_0) = s_0$.
+Secondly, we apply the assumption to the density in the sense that all the particles that start with the same initial state must end up in the same final state and vice-versa.
+That is, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$: the density must remains the same throughout the evolution.
 
-Now consider the integration $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
-The region $\Sigma$ will be mapped to a new region, but the fraction of the system found in both will be the same.
-If both the integral and $\rho_c$ have to remain during the evolution, then $\omega$ will also need to remain the same.
+Now consider the integral $\int_{\Sigma} \rho \omega(d\Sigma)$.
+Both the region and the density will be mapped in time to $\hat{\Sigma}$ and $\hat{\rho}$ respectively.
+The fraction of the system found in the new region will have to be the same as the one found in the old region.
+That is, $\int_{\Sigma} \rho \omega(d\Sigma) = \int_{\hat{\Sigma}} \rho \hat{\omega}(d\hat{\Sigma})$.
+Since both the integral and the density have to remain constant during the evolution, then $\omega$ will need to remain the same.
 That is, the areas in phase space must be mapped to areas of equal size and independent degrees of freedom must be mapped to independent degrees of freedom (or volumes would not be mapped to equal volumes).
@@
 
-\subsubsection{Motion equivalence}
+The argument can also be constructed through statistical concepts (i.e.
+deterministic and reversibility means conservation of variance), thermodynamic concepts (i.e.
+deterministic and reversibility means the only the state of the system is important for the evolution, the system is therefore isolated and must conserve energy) or informational theoretic consideration (i.e.
+under deterministic and reversible evolution the amount information does not change, information entropy has to be conserved).
 
-Having recovered classical Hamiltonian mechanics, we can further restrict the law of evolution with the following:
+\section{Kinematic equivalence}
 
-Motion equivalence assumption: the motion of the system (i.e. trajectories in physical space-time) is enough to recover its dynamics (i.e. evolutions in state space) and vice-versa.
+\textbf{Motion equivalence assumption:} \emph{the motion of the system (i.e. trajectories in physical space-time) is enough to recover its dynamics (i.e. evolutions in state space) and vice-versa.}
 
-If we focus on the parts, this means that for every evolution in phase space there should be one and only one trajectory.
+First, as before, we apply the assumption on each particle, which means that for every evolution in phase space there should be one and only one trajectory.
 Note that each space variable $x^i$ is a coordinate, a state variable that defines a unit.
@@
 
-In this case, that we call weak equivalence, $v^i(q^j, p_k)$ must be invertible at every $q^i$.
-Therefore Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
-Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ must be non-zero everywhere, and therefore must have the same sign which we take to be positive.
-This are exactly the cases where a Lagrangian can be constructed, and are exactly the cases where the Lagrangian leads to a unique solution.
+Let us call weak equivalence the notion that $v^i(q^j, p_k)$ must be invertible at every $q^i$.
+In this case, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
+Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ must be non-zero everywhere, and therefore must have the same sign which we take to be positive.
+In this case, and only in this case, we can construct a Lagrangian from a Hamiltonian.
+These are also exactly the cases where the Lagrangian, using the principle of minimal action, leads to a unique solution.
 
-If we look at the whole distribution, we have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
+Now we look at the whole distribution and how it can be expressed as a function of position and velocity.
+We have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
 \begin{equation}
@@
 
-Therefore
-
-\subsection{Fundamentality}
-
-Different notions of fundamentality, perfectly natural properties, naturalness, etc.
-
-Joint carving, perfectly natural properties.
-Ontological notion of fundamental.
-Clearly state why this is not "good" for this context.
-We are agnostic.
-One may use following points to argue whatever they want.
-
-Joint carving? What notion of fundamentality?
-
-\section{Argument for privileging Hamiltonian mechanics}
-
-\subsection{Relevant formal background from assumptions of physics framework}
-
-Hamiltonian mechanics makes fewer assumptions on the physical system being studied than Lagrangian mechanics.
-Lagrangian is a subset than Hamiltonian.
-Makes "amount of structure" more rigorous.
-Hamiltonian defines the relationship between conjugate momentum and velocity ($dq/dt = \partial H / \partial p$), so kinematic assumption constrains the Hamiltonian.
-
-Hamiltonian mechanics is "fundamental" because allows to show that densities are coordinate independent quantities and therefore physical quantities.
-Canonical pairs (position and conjugate momentum) allows a simpler expression/comparison of densities over space.
-That is why statistical mechanics works best on phase space (we have a uniform measure over q/p).
-
-Can characterize this argument as physically or epistemically privileging Hamiltonian mechanics
-
-\section{Interpreting Lagrangian mechanics}
-
-Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics.
-Curiel frequently appeals to naturalness and a distinction between natural vs.
-ad hoc/artificial constructions [figure out what his definition of ``naturalness" is, if he provides one].
-
-For instance, Curiel argues that Hamiltonian mechanics is less natural and thus less physically significant/privileged:  ``Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
-Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2)" (306).
-
-Response: symplectic geometry indeed does not care about the difference between q's and p's.
-Classical mechanics does.
-Misindentification of the math (symplectic geometry) with the physics (Classical Hamiltonian mechanics).
-
-Curiel's main point: the representation of a classical mechanical system in the Lagrangian framework is \textit{natural} in a way that its representation in a Hamiltonian formulation is not.
-The former representation depends on intrinsic structures of the classical mechanical system rather than an artificial construction (270).
-
-Part of our response: The action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
-It also holds when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
-Also shows that Lagrangian is ``unphysical" and depends on gauge.
-
-Upshot: can better understand Lagrangian mechanics as a special case of Hamiltonian mechanics.
-Explain how the geometrical understanding of the stationary action in extended phase space should meet any criterion of ``naturalness" or ``natural construction." Lagrangian mechanics is actually less natural because contingent on having a bijection between velocity and conjugate momentum.
-
-%
-
-\bibliography{bibliography}
-
 \end{document}

### 2020-10-27 Overviews/Classical mechanics/ClassicalMechanicsOverview.tex
Reviewed Classical Overview with Christine

@@
-\documentclass[10pt,twocolumn]{revtex4-1}
+\documentclass[10pt,twocolumn, nofootinbib]{revtex4-1}
 %\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
@@
 \usepackage{enumitem}
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks=true,
+	urlcolor=blue,
+	linkcolor=blue
+}
+\urlstyle{same}
 
@@
 
-\title{Classical mechanics and infinitesimal reducibility  \\
-	(Assumptions of Physics overview)}
+\title{Assumptions of Physics overview: \\
+	Classical mechanics and infinitesimal reducibility }
 \author{Gabriele Carcassi, Christine A. Aidala}
@@
 We briefly show how classical mechanics can be rederived and better understood as a consequence of three assumptions: infinitesimal reducibility, deterministic and reversible evolution, and kinematic equivalence.
-This work is an overview of some of the result of Assumptions of Physics (https://assumptionsofphysics.org), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+This work is an overview of some of the results of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
 \end{abstract}
@@
 The three assumptions lie on the left column.
-Each assumption lead to one or two key insights that progressively lead to the physical concepts in the middle column.
-Each of these is then mapped to its mathematical framework on the right.
+Each assumption leads to one or two key insights that progressively lead to the physical concepts in the middle column.
+Each of these is then mapped to its corresponding formal framework on the right.
 
@@
 
-\textbf{Infinitesimal reducibility assumption:} \emph{the state of the system is reducible to the state of its infinitesimal parts.} That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the state of its parts and so on.
+\textbf{Infinitesimal reducibility assumption:} \emph{the state of the system is reducible to the state of its infinitesimal parts.} That is, giving the state of the whole system is equivalent to giving the state of its parts, which in turn is equivalent to giving the state of its subparts and so on.
 
-Under this assumption, the state of the whole is a distribution over the state of the parts.
+Under this assumption, the state of the whole is a distribution over the states of the parts.
 More precisely, let $\mathcal{C}$ be the state space for the whole system.
@@
 
-The key concept is to keep track of the unit system so we need precise terminology.
-We call state variables a set of quantities $\xi^a$ (i.e.
-coordinates in manifold terminology) that fully identify a state.
-That is, we can write each state $s(\xi^a)$ as a function of the state variables.
+The key concept is to keep track of the unit system, so we need precise terminology.
+We call state variables a set of quantities $\xi^a$ that fully identify a state.\footnote{In mathematical terminology, these are the coordinates of the manifold.} That is, we can write each state $s(\xi^a)$ as a function of the state variables.
 We call coordinate $q \in \xi^a$ a particular variable that defines a unit.
@@
 
-We start with the simplest case, where one coordinates is sufficient to define the unit system, which means the following three conditions:
+We start with the simplest case, where one coordinate is sufficient to define the unit system, which means the following three conditions:
 \begin{enumerate}[noitemsep]
@@
 
-Now we show that there can only be one a single $k^i$.
+Now we show that there can only be one $k^i$.
 Suppose we change unit $\hat{q}=\hat{q}(q)$.
-Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
+Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^j\}$.
 We have $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$.
 Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must be equal to 1.
-Note that, since $\hat{q}$ depends only on $q$, $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$.
+Note that, since $\hat{q}$ depends only on $q$, $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^j}{\partial k^i} \right|$.
 Suppose there is only one variable.
@@
 This puts a constraint only on the determinant of the transformation.
-Suppose there are three or more variable.
-This constraint is not enough to recover the transformation uniquely and therefore $\hat q$  would not fully define the units for all other state variable.
-This means there must be two variable: $q$ and a single $k$.
+Suppose there are three or more variables.
+This constraint is not enough to recover the transformation uniquely and therefore $\hat q$  would not fully define the units for all other state variables.
+This means there must be two variables: $q$ and a single $k$.
 Coordinate independent areas and densities can only be defined on a two-dimensional manifold.
@@
 Now we generalize to multiple coordinates.
-We say two coordinates are independent if changing the units for one does not induce a change the units for the other.
+We say two coordinates are independent if changing the units for one does not induce a change of units for the other.
 Now suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$.
@@
 Now change the second coordinate $q^2$ in the same way while also fixing $k_1$.
-Then we find a correspond $k_2$.
+Then we find a corresponding $k_2$.
 We can proceed in the same way until we exhaust all coordinates, which must also mean that there are no state variables left.
 We find that $\mathcal{S}$ is $2n$-dimensional and the state variables are $\xi^a = \{ q^i, k_i \}$.
-We define an independent degree of freedom the space charted by a pair of such variables.
+We define an independent degree of freedom as the space charted by a pair of such variables.
 
@@
 Because degrees of freedom are, at least locally, independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
-This means the volume form is given by $\omega^n$.
-This cannot be degenerate, or some volumes would not contain states.
+This means the volume form is proportional to $\omega^n$.
+This cannot be degenerate (i.e.~it must be nonzero for each infinitesimal volume), or some volumes would not contain states.
 Therefore $\omega$ itself cannot be degenerate.
 As the degrees of freedom are independent, the number of states on a surface does not change if we translate it across independent degrees of freedom.
-If we imagine a parallelopiped, the integral over the surface must be zero (integral over opposite sides are equal and opposite).
-Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate and therefore is a symplectic manifold.
+If we imagine a parallelepiped, the integral over the surface must be zero (integrals over opposite sides are equal and opposite).
+Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate and therefore $\mathcal{S}$ is a symplectic manifold.
 By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
 
-Phase space (i.e. symplectic manifolds) is the only type of manifold that is able to support coordinate invariant distributions, which are required to describe an infinitesimally reducible system.
+Phase space (i.e.~symplectic manifold) is the only type of manifold that is able to support coordinate invariant distributions, which are required to describe an infinitesimally reducible system.
 
@@
 Secondly, we apply the assumption to the density in the sense that all the particles that start with the same initial state must end up in the same final state and vice-versa.
-That is, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$: the density must remains the same throughout the evolution.
+That is, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$: the density must remain the same throughout the evolution.
 
@@
 The evolution is a symplectomorphism and corresponds to Hamiltonian evolution.
-Intuitively, this is the inverse of Liousville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
+Intuitively, this is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
 
-The argument can also be constructed through statistical concepts (i.e.
-deterministic and reversibility means conservation of variance), thermodynamic concepts (i.e.
-deterministic and reversibility means the only the state of the system is important for the evolution, the system is therefore isolated and must conserve energy) or informational theoretic consideration (i.e.
-under deterministic and reversible evolution the amount information does not change, information entropy has to be conserved).
+The argument can also be constructed through statistical concepts (i.e.~determinism and reversibility means conservation of variance), thermodynamic concepts (i.e.~determinism and reversibility means only the state of the system is important for the evolution, the system is therefore isolated and must conserve energy) or information theoretic consideration (i.e.~under deterministic and reversible evolution the amount of information does not change, information entropy has to be conserved).
 
@@
 
-\textbf{Motion equivalence assumption:} \emph{the motion of the system (i.e. trajectories in physical space-time) is enough to recover its dynamics (i.e. evolutions in state space) and vice-versa.}
+\textbf{Kinematic equivalence assumption:} \emph{the motion of the system (i.e.~trajectories in physical space-time) is enough to recover its dynamics (i.e.~evolutions in state space) and vice-versa.}
 
-First, as before, we apply the assumption on each particle, which means that for every evolution in phase space there should be one and only one trajectory.
+First, as before, we apply the assumption to the particles, which means that for every evolution in phase space there should be one and only one trajectory.
 Note that each space variable $x^i$ is a coordinate, a state variable that defines a unit.
@@
 Since the trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
-If the equations of motions were such that $v^i=v^i(q^i)$, then kinematic equivalence would fail as the full trajectory would be determined only by $q^i$.
+If the equations of motion were such that $v^i=v^i(q^i)$, then kinematic equivalence would fail as the full trajectory would be determined only by $q^i$.
 So we must have $v^i=v^i(q^i, p_i)$.
 The relationship must be invertible or kinematic equivalence would fail.
-At any given time, then, we must the following relationship:
+At any given time, then, we must have the following relationship:
 \begin{equation}
@@
 x^i &= q^i \\
-v^i &= d_t x^i = v^i(q^j, p_k)
+v^j &= d_t x^j = v^j(q^i, p_k)
 \end{aligned}
@@
 
-Let us call weak equivalence the notion that $v^i(q^j, p_k)$ must be invertible at every $q^i$.
-In this case, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
-Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ must be non-zero everywhere, and therefore must have the same sign which we take to be positive.
-In this case, and only in this case, we can construct a Lagrangian from a Hamiltonian.
+Let us call weak equivalence the notion that $v^j(q^i, p_k)$ must be invertible at every $q^i$ and therefore we can write $p_k= p_k(x^i, v^j)$ as a function of position and velocity.
+In this case, the Jacobian matrix  $\frac{\partial v^j}{\partial p_k}$ must be invertible.
+Since $v^j = d_t q^j = \frac{\partial H}{\partial p_j}$, the Hessian $\frac{\partial^2 H}{\partial p_k \partial p_j}$ must be nonzero everywhere, and therefore must have the same sign which we take to be positive.
+In this case, and only in this case, we can construct a Lagrangian from a Hamiltonian:
+\begin{equation}
+	L(x^i, v^j) = v^k p_k (x^i, v^j) - H(q^i(x^i), p_k(x^i, v^j))
+\end{equation}
 These are also exactly the cases where the Lagrangian, using the principle of minimal action, leads to a unique solution.
@@
 If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only.
-Similar considerations will also hold for marginal distributions (i.e.
-distribution on a subset of the coordinates) which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
+Similar considerations will also hold for marginal distributions (i.e.~distributions on a subset of the coordinates) which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
 We set:
@@
 \end{equation}
-where $m$ is the unit conversion constant between velocity and conjugate momentum while $g_ij$ represents the linear dependency.
+where $m$ is the unit conversion constant between velocity and conjugate momentum while $g_{ij}$ represents the linear dependency.
 

### 2020-10-27 Overviews/Classical mechanics/ClassicalMechanicsOverview.tex
Added bib

@@
 	colorlinks=true,
+	citecolor=blue,
 	urlcolor=blue,
@@
 
-The overall argument can be summed up in the following diagram that can be used as a guide throughout this note.
+The overall argument (see \cite{AoPPhy1,AoPBook} for more details) can be summed up in the following diagram that can be used as a guide throughout this note.
 
@@
 
+\bibliography{bibliography}
+
 \end{document}

### 2020-11-01 Overviews/Classical mechanics/ClassicalMechanicsOverview.tex
Minor edits & Intro paragraph

@@
 We can proceed in the same way until we exhaust all coordinates, which must also mean that there are no state variables left.
-We find that $\mathcal{S}$ is $2n$-dimensional and the state variables are $\xi^a = \{ q^i, k_i \}$.
+We find that $\mathcal{S}$ is $2n$-dimensional, and the state variables are $\xi^a = \{ q^i, k_i \}$.
 We define an independent degree of freedom as the space charted by a pair of such variables.
@@
 
-Phase space (i.e.~symplectic manifold) is the only type of manifold that is able to support coordinate invariant distributions, which are required to describe an infinitesimally reducible system.
+Phase space (i.e.~a symplectic manifold) is the only type of manifold that is able to support coordinate invariant distributions, which are required to describe an infinitesimally reducible system.
 
@@
 
-The argument can also be constructed through statistical concepts (i.e.~determinism and reversibility means conservation of variance), thermodynamic concepts (i.e.~determinism and reversibility means only the state of the system is important for the evolution, the system is therefore isolated and must conserve energy) or information theoretic consideration (i.e.~under deterministic and reversible evolution the amount of information does not change, information entropy has to be conserved).
+The argument can also be constructed through statistical concepts (i.e.~determinism and reversibility means conservation of variance), thermodynamic concepts (i.e.~determinism and reversibility means only the state of the system is important for the evolution; the system is therefore isolated and must conserve energy) or information theoretic consideration (i.e.~under deterministic and reversible evolution the amount of information does not change, so information entropy has to be conserved).
 
@@
 First, as before, we apply the assumption to the particles, which means that for every evolution in phase space there should be one and only one trajectory.
-Note that each space variable $x^i$ is a coordinate, a state variable that defines a unit.
+Note that each space variable $x^i$ is a coordinate, i.e.
+a state variable that defines a unit.
 In fact, the trajectories can be fully described by those units and only those units.
 So we can say that $q^i=x^i$, each $q^i$ will be paired with a conjugate $p_i$ and each state $\{q^i, p_i\}$ will be mapped to one and only one trajectory.
-At each point $x^i$, then, must pass infinitely many trajectories, one for each combination of $\{p_i\}$.
+At each point $x^i$, then, infinitely many trajectories must pass, one for each combination of $\{p_i\}$.
 Since the trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.

### 2020-11-01 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex

@@
 
-\subsection{Define Lagrangian \& Hamiltonian mechanics}
+Philosophers have recently shown that we can interpret Lagrangian and Hamiltonian mechanics as theoretically equivalent in a precise mathematical sense.
+Within a subclass of classical mechanical systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh2017}{Barrett2018}.
+Categorical equivalence provides a natural interpretive criterion for theoretical equivalence because it yields mutual inter-translatability between theory formulations.
+At first glance, this equivalence seems to motivate the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain.
+This view would undermine prior debates that have privileged either Hamiltonian mechanics or Lagrangian mechanics on physical or metaphysical grounds \parencites[]{North2009}{Curiel2014}.
+Here, we will argue that there are indeed physical grounds for privileging Hamiltonian mechanics over Lagrangian mechanics, their categorical equivalence notwithstanding.
+Compared to Hamiltonian systems, an additional physical condition must be satisfied for a classical system to be Lagrangian.
+Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this conceptual difference has interpretive consequences even within the hyper-regular domain.
+Furthermore, the construction at the heart of this argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density, along with a novel interpretation of the principle of stationary action within the Hamiltonian formalism.
 
-Talk about disippative systems.
+Thesis: despite the mathematical equivalence of Hamiltonian and Lagrangian mechanics within the hyper-regular domain, the two frameworks have important differences.
+In particular, Hamiltonian systems arise when a smaller number of basic physical conditions are met.
+A Lagrangian system must satisfy a further physical condition.
+This shows that Hamiltonian mechanics is in a sense more fundamental.
+Defending this claim does not require commitments to fundamental structure or perfectly natural properties.
+
+Targets: Curiel's privileging of Lagrangian mechanics and argument that Hamiltonian mechanics is ad hoc; North's somewhat flawed privileging of Hamiltonian mechanics on the basis of it requiring less structure; a motivation stemming from categorical equivalence proofs for thinking that there are no physically significant differences between the Hamiltonian and Lagrangian frameworks (at least within the hyper regular domain).
+
+Goals: provide a gentle introduction to Assumptions of Physics framework, and show that it has real philosophical payoffs for interpreting physical theories, such as classical mechanics.
+
+\subsection{Define Classical Mechanics}
+
+Talk about dissipative systems.
 
@@
 Classical Hamiltonian systems satisfy the additional conditions of having deterministic and reversible evolutions.
-These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it contact manifold.
+These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it a contact manifold.
 Classical Lagrangian systems satisfy a further assumption: keeping track of the trajectories in physical space is equivalent to tracking the evolution of states in phase space.
@@
 This puts a constraint only on the determinant of the transformation.
-Suppose there are three or more variable.
-This constraint is not enough to recover the transformation uniquely and therefore $\hat q$  would not fully define the units for all other state variable.
-This means there must be two variable: $q$ and a single $k$.
-Coordinate independent areas and densities can only be defined on a two-dimensional manifold.
+Suppose there are three or more variables.
+This constraint is not enough to recover the transformation uniquely and therefore $\hat q$  would not fully define the units for all other state variables.
+This means there must be two variables: $q$ and a single $k$.
+So in this case, coordinate independent areas and densities can only be defined on a two-dimensional manifold.
 
@@
 Now change the second coordinate $q^2$ in the same way while also fixing $k_1$.
-Then we find a correspond $k_2$.
-In the end, we will find that $\mathcal{S}$ is $2n$-dimensional and the state variables are $\xi^a = \{ q^i, k_i \}$.
-We define an independent degree of freedom the space charted by a pair of such variables.
+Then we find a corresponding $k_2$.
+In the end, we will find that $\mathcal{S}$ is $2n$-dimensional, and the state variables are $\xi^a = \{ q^i, k_i \}$.
+We define an \textit{independent degree of freedom} as the space charted by a pair of such variables.
 
@@
 Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define the integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
-Because degrees of freedom are, at least locally, independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
+Because degrees of freedom are---at least locally---independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
 This means the volume form is given by $\omega^n$.
@@
 As the degrees of freedom are independent, the number of states on a surface does not change if we translate it across independent degrees of freedom.
-If we imagine a parallelopiped, the integral over the surface must be zero (integral over opposite sides are equal and opposite).
+If we imagine a parallelepiped, the integral over the surface must be zero (integral over opposite sides are equal and opposite).
 Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate and therefore is a symplectic manifold.
@@
 
-Deterministic and reversible evolution assumption: given the present state of the system, all future (determinism) and past (reversibility) states are uniquely identified.
+Deterministic and reversible evolution assumption: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.
 
-In our case, this goes to both the state of the whole and to the state of the parts.
+In our case, this applies to both the state of the whole and to the state of the parts.
 That is, given an initial state $s_0$ there will be one and only one evolution $s(t)$ such that $s(t_0) = t_0$.
-Moreover, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$ the density remains the same throughout the evolution.
-That is, all particles that start with the initial state must end up in the same final state and vice-versa.
+Moreover, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$, i.e.
+the density remains the same throughout the evolution.
+In other words, all particles that begin in the initial state must end up in the same final state and vice-versa.
 
-Now consider the integration $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
-The region $\Sigma$ will be mapped to a new region, but the fraction of the system found in both will be the same.
-If both the integral and $\rho_c$ have to remain during the evolution, then $\omega$ will also need to remain the same.
-That is, the areas in phase space must be mapped to areas of equal size and independent degrees of freedom must be mapped to independent degrees of freedom (or volumes would not be mapped to equal volumes).
+Now consider the integral $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
+The region $\Sigma$ will be mapped to a new region, but the fraction of the system found in both the old and new regions must be the same.
+If both the integral and $\rho_c$ have to remain constant during the evolution, then $\omega$ will also need to remain the same.
+That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (or volumes would not be mapped to equal volumes).
 The evolution is a symplectomorphism and corresponds to Hamiltonian evolution.
-Intuitively, this is the inverse of Liousville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
+Intuitively, this is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
 
@@
 
-Having recovered classical Hamiltonian mechanics, we can further restrict the law of evolution with the following:
+Having recovered classical Hamiltonian mechanics, the laws of evolution become further restricted when the following condition holds:
 
@@
 
-If we focus on the parts, this means that for every evolution in phase space there should be one and only one trajectory.
-Note that each space variable $x^i$ is a coordinate, a state variable that defines a unit.
-In fact, the trajectories can be fully described by those units and only those units.
+If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.
+Note that each space variable $x^i$ is a coordinate, i.e.
+a state variable that defines a unit.
+In fact, the trajectories can be fully described by these units and only these units.
 So we can say that $q^i=x^i$, each $q^i$ will be paired with a conjugate $p_i$ and each state $\{q^i, p_i\}$ will be mapped to one and only one trajectory.
-At each point $x^i$, then, must pass infinitely many trajectories, one for each combination of $\{p_i\}$.
+At each point $x^i$, then, infinitely many trajectories must pass, one for each combination of $\{p_i\}$.
 Since the trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
-If the equations of motions were such that $v^i=v^i(q^i)$, then kinematic equivalence would fail as the full trajectory would be determined only by $q^i$.
+If the equations of motions were such that $v^i=v^i(q^i)$, then motion equivalence would fail as the full trajectory would be determined only by the $q^i$.
 So we must have $v^i=v^i(q^i, p_i)$.
-The relationship must be invertible or kinematic equivalence would fail.
-At any given time, then, we must the following relationship:
+This relationship must be invertible or motion equivalence would fail.
+At any given time, then, we must have the following relationships:
 \begin{equation}
@@
 
-In this case, that we call weak equivalence, $v^i(q^j, p_k)$ must be invertible at every $q^i$.
-Therefore Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
-Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ must be non-zero everywhere, and therefore must have the same sign which we take to be positive.
-This are exactly the cases where a Lagrangian can be constructed, and are exactly the cases where the Lagrangian leads to a unique solution.
+In this case, which we call \textit{weak equivalence}, $v^i(q^j, p_k)$ must be invertible at every $q^i$.
+Therefore, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
+Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ must be non-zero everywhere, and therefore must have the same sign which we take to be positive.
+These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).
 
@@
 Similar considerations will also hold for marginal distributions (i.e.
-distribution on a subset of the coordinates) which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
+distributions on a subset of the coordinates) which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
 We set:
@@
 \begin{equation}
-v^i = d_t q^i = \frac{\partial H }{\partial P} = \frac{1}{m} g^{ij}(p_j - A_j) \\
+v^i = d_t q^i = \frac{\partial H }{\partial p_i} = \frac{1}{m} g^{ij}(p_j - A_j) \\
 \end{equation}

### 2020-11-02 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Edits to derivation & Brainstorm

@@
 
-\usepackage{amsmath, hyperref}
+\usepackage{amsmath, hyperref, enumerate}
 \usepackage{amsfonts}
@@
 \end{abstract}
+
+Current section ideas:
+
+\begin{enumerate}[1.)]
+
+\item Assumptions of physics framework section
+
+\begin{itemize}
+
+\item motivate everything upfront: going to show that Hamiltonian mechanics corresponds to three conditions being satisfied, while Lagrangian mechanics requires a fourth condition
+\item define the three assumptions, and give motivation
+\item figuring out which mathematical structures are required for systems that satisfy certain physical conditions
+\item go over the upshot of the philosophical argument, specifically how this avoids the worries about measuring the amount of structure that plagued North (i.e. Swanson and Halvorson paper)
+\item Then, give the derivation, as gentle and well-motivated as possible. But would be ideal if readers could skip or skim this section and still get the philosophical upshot, relatively early on.
+
+\end{itemize}
+
+\item Physical significance of invariant distributions
+
+\begin{itemize}
+
+\item explain how these provide a novel way of understanding things, one that we can physically privilege
+
+\item connect this with North on invariant structure, but disavow metaphysical commitments
+
+\end{itemize}
+
+\item How to interpret Lagrangian mechanics in light of Hamiltonian mechanics
+
+\begin{itemize}
+
+\item gauge transformation stuff
+\item go over the cool stationary action principle derivation/argument
+\item clarify the status of Riemannian metric structure in the Lagrangian formalism
+\item respond to Curiel's privileging of Lagrangian mechanics
+\item respond to Curiel's argument that Hamiltonian mechanics involves the imposition of an ad hoc condition in order to get the right kinematical constraints, whereas these fall out naturally from Lagrangian mechanics.
+
+\end{itemize}
+
+\item Considerations in light of categorical equivalence proofs
+
+\begin{itemize}
+
+\item  clarify that argument is not in tension with categorical equivalence within the hyper regular domain. make this precise.
+\item Respond to the worry that something must have gone amiss if we have a difference in the number of conditions required for the Hamiltonian vs.
+Lagrangian mechanics, yet still have categorical equivalence in the hyper-regular domain
+
+\end{itemize}
+
+\end{enumerate}
 
@@
 For instance, knowledge of the trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from its trajectory alone.
+Hence, the set of classical systems that satisfy this further physical condition is a proper subset of the classical systems that satisfy the assumptions for Hamiltonian mechanics.
 
@@
 
-Classical phase space is recovered under the:
+First, we will show that a classical phase space arises if and only if a system satisfies the following physical condition:
+
+\begin{quotation}
 Infinitesimal reducibility assumption: the state of the system is reducible to the state of its infinitesimal parts.
-That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the state of its parts and so on.
+That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the states of those parts, and so on.
+\end{quotation}
+
+\noindent
+This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
+[At least in a footnote: comment on how this characterization of infinitesimal particles relates to standard philosophy of physics interpretive disputes over the status of point particles in classical mechanics.
+See Butterfield Pointillisme article; Butterfield characterizes particles as extensionless.]
 
 Let $\mathcal{C}$ be the state space for the whole system.
-We call particle the limit of recursive subdivision.
-Let $\mathcal{S}$ be the state space for the particles.
-Then for each $c \in \mathcal{C}$ we can find one and only one $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of its parts.
-That is, the state of the whole is a distribution over the infinitesimal parts.
-If $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space.
+Any point $c \in \mathcal{C}$ represents a particular state of the whole system.
+We define a classical \textit{particle} as the limit of recursive subdivision.
+A classical system then consists of a collection of classical particles.
+Let $\mathcal{S}$ be the state space for the particles [all of the particles constituting the whole system?
+or for each particle individually?].
+We assume that this state space $\mathcal{S}$ is a manifold.
+Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
+That is, the state of the whole system is a distribution over its infinitesimal parts.
+If $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space $\mathcal{S}$.
+%[is region U in the state space C or S. I assume in S.].
 
-The critical issue is that, on a manifold, $\rho_c$ must transform both as a scalar function (the value must depend on the point and not on the coordinates) and a density.
-Classical phase space (i.e.
-a symplectic manifold) is the only space that have this structure.
-Let's see why.
+On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function, i.e.
+it must depend on the point [the point $c $?
+] and not on the coordinates.
+This is because the distribution is a function only of the states, rather than of the particular choice of state variables, i.e.
+coordinates.
+[Explain why these two physical conditions hold for the distribution.
+Also, explain what it means to transform as a density.] We will now show that classical phase space (i.e.
+a symplectic manifold) is the only space that satisfies these constraints on the distribution.
 
-We call state variables a set of quantities $\xi^a$ (i.e.
-coordinates in manifold terminology) that fully identify a state.
-We call coordinate $q \in \xi^a$ a particular variable that defines a unit.
-Suppose the particle state space is such that one coordinate is sufficient to define the unit system.
-That is, $\xi^a = \{ q, k^i \}$.
-We need to make sure that we can arbitrarily change the unit without changing $\rho_c$.
-We also have to make sure that $q$ fully defines the units for all state variables.
-How many $k^i$ can we have?
+We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state (in manifold terminology, these are known as coordinates).
+We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
+For instance, $q$ might be position, defining units of meters.
+Starting with the simplest case, suppose that one coordinate suffices to define the unit system for the particle state space.
+Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number.
+We will show that, in fact, there can only be a single $k^i$ coordinate in this case, conjugate to the single $q$ coordinate.
+Our derivation assumes---on physical grounds---that we can (i) arbitrarily change the unit defined by $q$ without changing the density $\rho_c$ and that (ii) $q$ fully defines the units for all state variables.
+%[Give a concrete example of a state variable defining a unit, along with physical motivation for why state variables play the role of defining units (since we are treating this as a physically grounded assumption).]
+% How many $k^i$ can we have?
 
-Suppose we change unit $\hat{q}=\hat{q}(q)$.
+[We need to introduce the definition of state variable evolution $s$, since we appeal to it below.
+I take it that the definition we are working with is that future states are labeled by the original state variable values (i.e.
+transported state variables)?
+Section IV.C of long paper.]
+
+Now, suppose that we change the units by transforming the coordinate $q$ to a new coordinate $\hat{q}$ that is a function of the old coordinate: $\hat{q}=\hat{q}(q)$.
 Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
@@
 Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must be equal to 1.
-Note that, since $\hat{q}$ depends only on $q$, $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$.
-Suppose there is only one variable.
-Then we would have $\left|\frac{\partial \hat q}{\partial q} \right| = 1$.
+Suppose there is only one state variable, so that there are no $k^i$ variables.
+Then we would have $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right| = 1$.
 But this would mean the unit change cannot be arbitrary.
-Therefore we must have two or more variables and therefore $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$.
-This puts a constraint only on the determinant of the transformation.
+Therefore we must have two or more variables, so that there is at least one $k^i$ variable.
+The requirement that the Jacobian determinant equals one places a constraint on how the state variables $k^i$ must change to compensate the change from $q$ to $\hat q$.
+Note that, since $\hat{q}$ depends only on $q$, the Jacobian is an upper triangular block matrix, so $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$ [it might be worthwhile to spell this out in more detail].
+For the Jacobian determinant to equal one, we must therefore have $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$.
+This places a constraint only on the determinant of the transformation.
 Suppose there are three or more variables.
-This constraint is not enough to recover the transformation uniquely and therefore $\hat q$  would not fully define the units for all other state variables.
-This means there must be two variables: $q$ and a single $k$.
-So in this case, coordinate independent areas and densities can only be defined on a two-dimensional manifold.
+This constraint would then be insufficient to recover the transformation uniquely, and therefore $\hat q$  would not fully define the units for all other state variables (contrary to assumption).
+This means there must be exactly two state variables: $q$ and a single $k$.
+So in this case, coordinate-independent areas and densities can only be defined on a two-dimensional manifold.
 
-To generalize, we say two coordinates are independent if changing the units for one does not change the units for the other.
+To generalize, we say two coordinates are \textit{independent} if changing the units for one does not change the units for the other.
 Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$.
@@
 Invariant densities can only be assigned to infinitesimal areas.
-Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define the integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
+Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$ [clarify which phase space $\Sigma$ belongs to].
 Because degrees of freedom are---at least locally---independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
 This means the volume form is given by $\omega^n$.
-This cannot be degenerate, or some volumes would not contain states.
+This cannot be degenerate, or some volumes would not contain states [motivate this claim or].
 Therefore $\omega$ itself cannot be degenerate.
 As the degrees of freedom are independent, the number of states on a surface does not change if we translate it across independent degrees of freedom.
-If we imagine a parallelepiped, the integral over the surface must be zero (integral over opposite sides are equal and opposite).
-Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate and therefore is a symplectic manifold.
+If we imagine a parallelepiped, the integral over the surface must be zero (the integrals over opposite sides are equal and opposite).
+Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold.
 By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
 
-The conclusion is that, under the infinitesimal reducibility assumption, the state space of the system is a coordinate invariant distribution over the state space of the infinitesimal parts.
-If the state space of the particles is a manifold, then it is a symplectic manifold, which is the only type of manifold that supports coordinate invariant distributions.
+The conclusion is that, under the infinitesimal reducibility assumption, the state space of the system is a coordinate-invariant distribution over the state space of the infinitesimal parts [what does it mean for the state space $\mathcal{C} $ to itself be a distribution ?].
+If the state space of the particles is a manifold, then it is a symplectic manifold, which is the only type of manifold that supports coordinate-invariant distributions [do we have a reference to substantiate the ``only" claim?
+Has the proof above really shown that this is the only kind of manifold that would support a coordinate-invariant distribution].
 
@@
 
+Next, we assume that the classical systems of interest are deterministic and reversible. This amounts to satisfying the following joint condition:
+
+\begin{quotation}
+
 Deterministic and reversible evolution assumption: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.
 
-In our case, this applies to both the state of the whole and to the state of the parts.
-That is, given an initial state $s_0$ there will be one and only one evolution $s(t)$ such that $s(t_0) = t_0$.
-Moreover, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$, i.e.
+\end{quotation}
+
+[Should we footnote that these two assumptions leave out some classical systems, e.g. Norton's dome, Earman's space invaders?]
+
+In our case, this applies to both the state of the whole system and to the state of its parts.
+That is, given an initial state $s_0$ there will be one and only one evolution $s(t)$ such that $s(t_0) = t_0$ [is $s$ an evolution that transports the state variables, i.e.
+labeling future states by the original state variables so that the initial time is transported to itself?
+I think we need to explain more about how evolutions are defined].
+Moreover, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle [what exactly does `initial particle' mean?
+Do we mean the initial state of the composite system or just a particular particle?] at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$, i.e.
 the density remains the same throughout the evolution.
-In other words, all particles that begin in the initial state must end up in the same final state and vice-versa.
+In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa [could two different particles occupy the same initial state?].
 
@@
 The region $\Sigma$ will be mapped to a new region, but the fraction of the system found in both the old and new regions must be the same.
-If both the integral and $\rho_c$ have to remain constant during the evolution, then $\omega$ will also need to remain the same.
-That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (or volumes would not be mapped to equal volumes).
+Since both the integral and $\rho_c$ must remain constant during the evolution, the two-form $\omega$ must also remain the same.
+That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (otherwise, volumes would not be mapped to equal volumes, which would violate determinism).
 The evolution is a symplectomorphism and corresponds to Hamiltonian evolution.
-Intuitively, this is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
+Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
 
@@
 where V is another arbitrary function. We recognize this as the Hamiltonian for massive particles under potential forces.
+
+[Connect this with Curiel equation 6.2 for the Hamiltonian.]
 

### 2020-11-05 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
Version submitted to PhysLetters

--- 2020-03-31 Articles/2019-QuantumComposition/Submissions/NatureComm/supplementaryinfo.tex
@@
 		the space.
-The exception is proposition \ref{prop_fundProj} for which the prove the finite case by induction and then show that it holds in the limit.
+The exception is proposition \ref{prop_fundProj} for which we prove the finite case by induction and then show that it holds in the limit.
 This would not work in the non-separable case, since the basis would not be countable.
@@
 conjugate-linear).
+This is not new, it is essentially Wigner's theorem, but the proof we offer is insightful as it clearly shows the connection between the construction of the map and the choice of gauge.
 	\end{remark}
@@
 	\begin{remark}
-		The fact that the proposition identifies either a linear map or an anti-linear (i.e.~conjugate-linear)  has a clear physical interpretation.
-Since, for a Hilbert space, the conjugate vector space is equivalent to the dual space, a conjugate-linear map is one that preserves the inner product but maps ket vectors into bra vectors.
+		The fact that the proposition identifies either a linear map or an anti-linear (i.e.~conjugate-linear) corresponds, in physics terms, to a choice of convention.
+As analogies: a change in metric signature in relativity would change the mathematical space but not the physics; in classical phase-space, a change in signature of the symplectic form would change the mathematical space, but not the physics it represents.
+These choices are widely recognized as a matter of personal preference.
+
+		In simple terms, for a Hilbert space, the conjugate
+                vector space is equivalent to the dual space, so we
+                could equivalently choose one or the other. For
+                example, Schr\"odinger in the papers in which he
+                introduces the Schr\"odinger equation, he writes it
+                with both signs, as the choice of sign of the imaginary part of the wave function is arbitrary: one sign refers to
+                the Hilbert space, the other to the dual space, namely
+                to the complex-conjugate wave function \cite{sch}.
+So we can think of the anti-linear map as one that preserves the inner product but maps ket vectors into bra vectors.
 Looking ahead, the above result does not exclude a composition map similar to the tensor product, but that maps the kets of one or both subsystems into bras in the composite system.
+This would only make the representation of the composite physical system more complicated, as we need to keep track of the different conventions in the different subspaces.
+Therefore, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear.
+With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.
 
-		Such a conjugate-linear map is physically indistinguishable from a map that maps kets into kets: the physics is only contained in the probabilities which are the square moduli of inner products.
-The idea is that, without changing the physics, we can always mathematically redefine the second space so that the resulting map is linear.
-With this in mind, we will assume that the map between the spaces is linear, which will in turn lead to identifying the tensor product as a unique composition map.
+Another way to look at this is that Hermitian operators, and therefore all 
+ the physics, are invariant under an anti-linear transformation.
+ In contrast, anti-Hermitian operators will change sign. This 
+ changes the connection between the generators and the generated 
+ transformations (i.e.~while $A$ generates $exp(\frac{Ab}{\imath\hbar})$ 
+ on one space, the mapped $A$ will generate 
+ $exp(-\frac{Ab}{\imath\hbar})$ in the mapped space). Note that the 
+ choice of whether to put the minus or not is arbitrary as long as one is consistent across all generators and transformations.
+ Similarly, we typically define $[A, B] = AB - BA$ but we could have 
+ alternatively chosen $[A, B] = BA - AB$. The anti-linear map is simply 
+ a change of that convention.
 
@@
 	\begin{prop}[Bilinearity, H2]\label{prop_bilinearity}
-		Let $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ be a map such that, for all $(a,b) \in \mathcal{A} \times \mathcal{B}$, we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$.
-Then $m$ must be bilinear.
+		Given $M$ in \ref{def_comp}, if we can find an $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ such that for all $(a,b) \in \mathcal{A} \times \mathcal{B}$, we have $\pj{m(a,b)} = M(\pj{a}, \pj{b})$, then $m$ must be bilinear.
 That is:
@@
 		quantum theory, (North Holland, 1982).
+		\bibitem{sch}E. Schr\"odinger, Annalen der Physik {\bf 102}, 81 (1926); English translation in E. Schr\"odinger, {\em Collected papers on Wave Mechanics} (Blackie \& Son, London, 1928).
 	\end{references}

### 2020-11-05 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
Version submitted to PhysLetters

--- 2020-04-12 Articles/2019-QuantumComposition/Submissions/NatureComm/tensor.natcom.tex
@@
 \title{The four postulates of quantum mechanics are three}
-	\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^{2,*}$ and Christine A. Aidala$^1$
+\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^{2}$ and Christine A. Aidala$^1$
 }\affiliation{\vbox{1.~Physics Department, University of Michigan, 450 Church Street,
@@
 \begin{figure}[ht]
-		%\epsfxsize=1.\hsize\leavevmode\epsffile{fig.eps}
-		\includegraphics[width=\linewidth, trim={0.2in 5.8in 2.5in 0.1in}, clip=true]{fig.eps}
+\epsfxsize=1.\hsize\leavevmode\epsffile{fig.eps}
+%\includegraphics[width=\linewidth, trim={0.2in 5.8in 2.5in 0.1in}, clip=true]{fig.eps}
 \caption{Schematic depiction of the logical implications used
@@
   clarify the assumptions that it contains. We will use the following
-		definition for a quantum system\togli{$\stackon[1pt]={\mbox{\tiny
+  definition for a quantum
+  system\togli{$\stackon[1pt]={\mbox{\tiny
         def}}$}$\stackrel{\mbox{\tiny def}}=${\em ``a quantum degree
@@
     property''}. \commentg{We may have to revise to be more clear.
-			Where is this used?} } As stated above, we will limit ourselves to
-	kinematically-independent systems, where all state vectors $|\psi\>$
-	in the system's Hilbert space $\cal H$ describe a valid state, {\em
-		unconditioned on anything else}. In particular, we note that
-	restrictions due to superselection rules arise either from practical
-	(not fundamental) limitations on the actions of the experimenter
-	\cite{susskind,zanardi,zanardilloyd} or from the use of ill-defined
-	quantum systems. For example, in situations where indistinguishability
-	plays a role (e.g.~in QFT), one cannot consider an electron as a
-	quantum system: in this case the quantum system is the field. The
-	electron is an excitation of the field, {\em not} a (sub)system. We
-	call this condition ``preparation independence''.  [We emphasize that
-	the kinematic independence is inequivalent to dynamical independence
-	(or isolation).  Indeed if two systems interact, their interaction may
-	lead to dynamical restrictions in the state spaces. We will not
-	consider dynamical evolution in this paper, which is contained in
+    Where is this used?} } As discussed, we limit ourselves to
+kinemati\-cal\-ly-inde\-pen\-dent systems, where all state vectors
+$|\psi\>$ in the system's Hilbert space $\cal H$ describe a valid
+state, {\em unconditioned on anything else}. In particular, we note
+that restrictions due to superselection rules arise either from
+practical (not fundamental) limitations on the actions of the
+experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
+ill-defined quantum systems. For example, in situations where
+indistinguishability plays a role (e.g.~in QFT), one cannot consider
+an electron as a quantum system: in this case the quantum system is
+the field. The electron is an excitation of the field, {\em not} a
+(sub)system. We call this condition ``preparation independence''.  [We
+emphasize that the kinematic independence is inequivalent to dynamical
+independence (or isolation).  Indeed if two systems interact, their
+interaction may lead to dynamical restrictions in the state spaces.
+Here we will not consider dynamical evolution, which is contained in
 postulate (d).]
@@
 corresponds to the map $m$ mapping kets of one or both subsystems into
-	bras before outputting the composite system (all these maps are
+bras before outputting the composite system. All these maps are
 trivially physically equivalent, since the dual-space description of a
 quantum system through bras is equivalent to the description using
-	kets). We can now repeat the same analysis for the second argument of
-	$m$ to conclude that it is a bilinear map, condition (H2).
+kets, or in other words, the wavefunction and its complex-conjugate
+contain identical information (for example, in his first papers
+Schr\"odinger used both signs in his equation: effectively writing
+{\em two} equivalent equations with complex-conjugate solutions
+\cite{sch}). We can now repeat the same analysis for the second
+argument of $m$ to conclude that it is a bilinear map, condition (H2).
 
@@
 (a), the measurement postulate (b) and the definitions of composite
-	and independent systems. In the methods section we prove that these
-	three conditions imply that the (up to now unspecified) composition
-	rule $m$ is the tensor product. More precisely, given a total,
-	span-surjective, bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$
-	that preserves the square of the inner product, we find that $\cal C $
-	is equivalent to $\cal A \otimes \cal B $ and that $m=\otimes$.
+and independent systems. We now prove that these three conditions
+imply that the (up to now unspecified) composition rule $m$ is the
+tensor product. More precisely, given a total, span-surjective,
+bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ that preserves the
+square of the inner product, we find that $\cal C $ is equivalent to
+$\cal A \otimes \cal B $ and that $m=\otimes$.
 
-	A few comments on the proof (given below): it is based on the universal
-	property of the tensor product, which uniquely characterizes it. First
-	we show that the bilinear map $m$ maps subsystems' bases into the
-	composite system basis. We also know that there exists a tensor
-	product map $ {T}=\otimes$ that can compose the vectors in $\cal A$
-	and $\cal B$.  Then we use the universal property: since $m$ is a
-	bilinear map, we are assured that there exists a unique $\hat m$ such
-	that $\hat m \circ T=m$. Since we show that $\hat m$ is an
-	isomorphism, then $\hat{m}$ bijectively maps vectors in $\cal C$ onto
-	vectors in the tensor product space. Namely $m={T}=\otimes$.
-
-	We conclude with some general comments. The tensor product structure
-	of quantum systems is not absolute, but depends on the observables
-	that are accessible \cite{zanardi,zanardilloyd}. This is due to the
-	fact that an agent that has access to a set of observables will define
-	quantum systems differently from an agent that has access to a
-	different set of observables. Where one agent sees a single system, an
-	agent that has access to less refined observables (and is then limited
-	by some superselection rules) can consider the same system as composed
-	of multiple subsystems. A typical example \cite{tellerbook} comes from
-	quantum field theory. It is customary in basically all quantum optics
-	literature to treat different modes of the radiation field (e.g.~the
-	output of two lasers) as independent systems composed through the
-	tensor product.  Clearly the electromagnetic field is a single system
-	and an agent who is able to access an optical mode that is a linear
-	combination of the two will give a quantum description for it that
-	cannot easily accommodate tensor products. Similarly, an agent can
-	consider two electrons as two systems, joined with the tensor product,
-	whenever they are distinguishable for all practical purposes (e.g.~the
-	electrons are in widely separated physical locations). Yet, in
-	principle, electrons are just excitations of a field, and the `true'
-	quantum system is the field, not the single electrons
-	\cite{teller,tellerbook}.  So, in quantum field theory, the quantum
-	systems that should be joined through tensor products are the
-	different fields and {\em not} the particles, which are just
-	excitations (states) of the fields. In the words of Teller
-	(\cite{tellerbook}, pg.22), tensor products can be safely used only if
-	there is a ``primitive thisness'', which is captured in the definition
-	of system.
-
-	It has been pointed out before that the quantum postulates are
-	redundant: in \cite{masanes,hartle} it was shown that the measurement
-	postulate (b) can be derived from the others (a), (c), (d). Here
-	instead we have shown how the tensor product postulate (c) can be
-	logically derived from the state postulate (a), the measurement
-	postulate (b) and a reasonable definition of independent systems, and
-	we have described the logical relations among them.  Of course, we do
-	not claim that this is the {\em only} way to obtain the tensor product
-	postulate from the others.
-
-	\vskip 1\baselineskip
-	{\par\noindent\bf Methods}
-
-	Proof of the statement: {\em given a total, span-surjective, bilinear
-		map $m:{\cal A}\times{\cal B}\to{\cal C}$ that preserves the square
-		of the inner product, we find that $\cal C $ is equivalent to $\cal
-		A \otimes \cal B $ and that $m=\otimes$.}
-
-	Step 1: the bases of the component systems are mapped to a basis of
+Proof. Step 1: the bases of the component systems are mapped to a basis of
 the composite system. Because of totality property (H1) and because
@@
 
+\vskip 1\baselineskip A few comments on the proof: it is based on the
+universal property of the tensor product, which uniquely characterizes
+it. First we show that the bilinear map $m$ maps subsystems' bases
+into the composite system basis. We also know that there exists a
+tensor product map $ {T}=\otimes$ that can compose the vectors in
+$\cal A$ and $\cal B$.  Then we use the universal property: since $m$
+is a bilinear map, we are assured that there exists a unique $\hat m$
+such that $\hat m \circ T=m$. Since we show that $\hat m$ is an
+isomorphism, then $\hat{m}$ bijectively maps vectors in $\cal C$ onto
+vectors in the tensor product space. Namely $m={T}=\otimes$.
+
+We conclude with some general comments. The tensor product structure
+of quantum systems is not absolute, but depends on the observables
+that are accessible \cite{zanardi,zanardilloyd}. This is due to the
+fact that an agent that has access to a set of observables will define
+quantum systems differently from an agent that has access to a
+different set of observables. Where one agent sees a single system, an
+agent that has access to less refined observables (and is then limited
+by some superselection rules) can consider the same system as composed
+of multiple subsystems. A typical example \cite{tellerbook} comes from
+quantum field theory. It is customary in basically all quantum optics
+literature to treat different modes of the radiation field (e.g.~the
+output of two lasers) as independent systems composed through the
+tensor product.  Clearly the electromagnetic field is a single system
+and an agent who is able to access an optical mode that is a linear
+combination of the two will give a quantum description for it that
+cannot easily accommodate tensor products. Similarly, an agent can
+consider two electrons as two systems, joined with the tensor product,
+whenever they are distinguishable for all practical purposes (e.g.~the
+electrons are in widely separated physical locations). Yet, in
+principle, electrons are just excitations of a field, and the `true'
+quantum system is the field, not the single electrons
+\cite{teller,tellerbook}.  So, in quantum field theory, the quantum
+systems that should be joined through tensor products are the
+different fields and {\em not} the particles, which are just
+excitations (states) of the fields. In the words of Teller
+(\cite{tellerbook}, pg.22), tensor products can be safely used only if
+there is a ``primitive thisness'', which is captured in the definition
+of system.
+
+It has been pointed out before that the quantum postulates are
+redundant: in \cite{masanes,hartle} it was shown that the measurement
+postulate (b) can be derived from the others (a), (c), (d). Here
+instead we have shown how the tensor product postulate (c) can be
+logically derived from the state postulate (a), the measurement
+postulate (b) and a reasonable definition of independent systems, and
+we have described the logical relations among them.  Of course, we do
+not claim that this is the {\em only} way to obtain the tensor product
+postulate from the others.
+
+{\it Acknowledgements:} L.M. acknowledges useful discussions with
+M.~Ozawa, P.~Zanardi, S.~Lloyd, D.~Zeh, G.~Auletta, A.~Aldeni and
+funding from the Attract project through the Eu Horizon 2020 research
+and innovation programme under grant agreement No 777222.
+G.C.
+and C.A.A.
+would like to thank M.
+J.
+Greenfield for reviewing the mathematical details and acknowledge funding from the MCubed program of the University
+	of Michigan.
+
 \begin{references}
@@
 \bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
+\bibitem{sch}E. Schr\"odinger, Annalen der Physik {\bf 102}, 81 (1926); English translation in E. Schr\"odinger, {\em Collected papers on Wave Mechanics} (Blackie \& Son, London, 1928).
 \bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
@@
 \end{references}
-
-	\vskip 1\baselineskip
-	{\bf Acknowledgements:} L.M. acknowledges useful discussions with
-	M.~Ozawa, P.~Zanardi, S.~Lloyd, D.~Zeh, G.~Auletta, A.~Aldeni and
-	funding from the Attract project through the Eu Horizon 2020 research
-	and innovation programme under grant agreement No 777222. G.C. and
-	C.A.A. acknowledge funding from the MCubed program of the University
-	of Michigan.
-
-	* Corresponding Author: maccone@unipv.it.
 

### 2020-11-05 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
Changes based on referees comments

@@
 
-	The following section needs to distinguish the projective space from
-	the Hilbert space itself. As this is seldom done in quantum mechanics,
-	we review some concepts in that context and introduce the notation
-	that we will be using. If $X$ is a Hilbert space, we denote $\pj{X}$
+	The core idea is that the probability space constrains what the state space for the composite system can be.
+Therefore we must develop a precise map between events in probability and their correspondents in terms of Hilbert space.
+Conceptually, the event $X = x_0$, the observable $X$ is equal to the value $x_0$, will correspond to the subspace spanned by all eigenstates of $X$ with eigenvalue $x_0$.
+The event $\psi$, the system was prepared in state $\psi$, will correspond to the ray (one-dimensional complex subspace) corresponding to the $\psi$ vector.
+Therefore, in general, all events in probability will corresponds to subspaces (of different dimensionality) of the Hilbder space.
+Projective spaces are the right tool to keep track of subspaces.
+The proof, then, consists in establishing what are the correct definitions in the space of events, map those into elements of the projective space and then, from the projective space, construct a map on the vector space directly.
+
+	Let us establish, then, the notation we will be using to distinguish the projective space from
+	the Hilbert space itself. If $X$ is a Hilbert space, we denote $\pj{X}$
 	the projective space. The projective space is mathematically
@@
 	of rays in the projective space, which we note as $\pj{Sp(v_1, ...,
-		v_n)}$. Given $v,w \in X$, we can write $P(v|w) = \frac{|\< v | w
+		v_n)}$. Geometrically, this can be thought as the smallest hyper-plane that contains all vectors. Given $v,w \in X$, we can write $P(v|w) = \frac{|\< v | w
 		\>|^2}{\< v | v \>\< w | w \>}$ which corresponds to the probability
@@
 	one can write $P(\pj{v} | \pj{w})\equiv P(v|w)$ as a function of the
-	rays.
+	rays. Geometrically, this corresponds to the angle between the two complex planes identified by the two vectors.
 
@@
 		(Born rule).\end{post}
+
+	{Given two events $\pj{a}$ and $\pj{b}$, for example $X > x_1$ and $X < x_2$, their conjunction $\pj{a} \wedge \pj{b}$ is the event where both are true, $ x_1 < X < x_2$ in the example.
+In terms of our Hilbert spaces, both $\pj{a}$ and $\pj{b}$ corresponds to subspaces and $\pj{a} \wedge \pj{b}$ is exactly the intersection of the two, which is also a subspace.
+We should not confuse $\pj{a} \wedge \pj{b}$ with $\pj{a} | \pj{b}$: the first refers to either
+		preparation or measurement of both systems  in the respective state, while the second corresponds to preparing one system in one state and measuring the
+		other system in the other state \cite{cox}.}
 
@@
 In other words, $\pj{a} \wedge \pj{b}$ and $M(\pj{a},\pj{b})$ correspond to the same event in probability space\footnote{We will end up proving that the map $M$ leads to the tensor product.}.
-			\item Every preparation of the composite gives a non-trivial measurement on the components.
+			\item For every preparation of the composite, local projective measurements must have at least one
+			outcome with non-zero probability.
 Formally, for every $\pj{c} \in \pj{\mathcal{C}}$, we can find at least $\pj{a} \in \pj{\mathcal{A}}$ and $\pj{b} \in \pj{\mathcal{B}}$ such that $P(\pj{a} \wedge \pj{b}|\pj{c})\neq 0$.
 		\end{enumerate}
-		Requirement 1 ensures that the composite system contains all the properties of the components.
+		It is important to understand that these requirements are necessary.
+Requirement 1 ensures that the composite system is well defined at least when the components are prepared independently.
+Conceptually, this ensures that the composite system contains all the properties of the components.
+Note that superselection rules or other restriction may prevent the independent preparation of all possible pairs (e.g.~two fermions cannot be jointly prepared in the same state).
+The tensor product is recovered only all pairs are compatible.
 Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components.

### 2020-11-05 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Fixing some math notation

@@
 In our case, this applies to both the state of the whole system and to the state of its parts.
-That is, given an initial state $s_0$ there will be one and only one evolution $s(t)$ such that $s(t_0) = t_0$ [is $s$ an evolution that transports the state variables, i.e.
+Let $\lambda : \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
+Under the assumption, this will be uniquely identified by the initial state $s_0 = \lambda(t_0)$ at the initial time $t_0$.
+[is $s$ an evolution that transports the state variables, i.e.
 labeling future states by the original state variables so that the initial time is transported to itself?
 I think we need to explain more about how evolutions are defined].
-Moreover, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle [what exactly does `initial particle' mean?
-Do we mean the initial state of the composite system or just a particular particle?] at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$, i.e.
+Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial particle [what exactly does `initial particle' mean?
+Do we mean the initial state of the composite system or just a particular particle?] at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e.
 the density remains the same throughout the evolution.
@@
 \begin{equation}
-|J| = \begin{bmatrix}
-\frac{\partial x^i}{\partial q^j} & \frac{\partial v^i}{\partial q^j} \\
-\frac{\partial x^i}{\partial p_j} & \frac{\partial v^i}{\partial p_j}
-\end{bmatrix}
-= \begin{bmatrix}
-\delta^i_j & \frac{\partial v^i}{\partial q^j} \\
-0 & \frac{\partial v^i}{\partial p_j}
-\end{bmatrix}
+\begin{aligned}
+|J| &= \begin{vmatrix}
+\frac{\partial x^i}{\partial q^j} & \frac{\partial x^i}{\partial p_j} \\
+\frac{\partial v^i}{\partial q^j} & \frac{\partial v^i}{\partial p_j}
+\end{vmatrix}
+= \begin{vmatrix}
+\delta^i_j & 0 \\
+\frac{\partial v^i}{\partial q^j} & \frac{\partial v^i}{\partial p_j}
+\end{vmatrix} \\
+&= \left|\delta^i_j\right| \left|\frac{\partial v^i}{\partial p_j}\right| - \left|0\right| \left|\frac{\partial v^i}{\partial q^j}\right|
 = \left|\frac{\partial v^i}{\partial p_j}\right|.
+\end{aligned}
 \end{equation}

### 2020-11-07 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex

--- 2020-11-02 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 
-\item motivate everything upfront: going to show that Hamiltonian mechanics corresponds to three conditions being satisfied, while Lagrangian mechanics requires a fourth condition
-\item define the three assumptions, and give motivation
+\item motivate everything upfront: going to show that Hamiltonian mechanics corresponds to two conditions being satisfied, while Lagrangian mechanics requires a three condition
+\item define the two assumptions, and give motivation
 \item figuring out which mathematical structures are required for systems that satisfy certain physical conditions

### 2020-11-07 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex

@@
 insufficient to single out the tensor product, as other linear
-products of linear spaces exist, such as the direct product, the
-topological product or the direct sum of vector spaces, which are used
+products of linear spaces exist, such as the direct product, the exterior/wedge roduct or the direct sum of vector spaces, which is used
 	in classical mechanics to combine state spaces of linear systems.
+These are all maps from linear spaces to linear spaces but they differ in how the linearity of one is mapped to the linearity of the other.\footnote{For example, in the tensor product $a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct product $a \times (b+c) = a \times b + 0 \otimes c$ where $0$ is the zero vector.
+Also, in the tensor product $r * (a \otimes b) = (r*a) \otimes b = a \otimes (r*b)$ while in the direct product $r * (a \times b) = r*a \times r*b$.}
+
 	This belief may have arisen from the seminal book of Dirac
@@
 	from the natural definition of a composite system as the set of two
-(or more) quantum systems, in the sense that the composite system is
-made of system $A$ {\em and} (joined with) system $B$, namely a set
-whose elements are the two systems and nothing else. We will focus on
+	(or more) quantum systems. The composite system is therefore
+	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else},
+	it's state being fully described by all observables and only the observables
+	that fully describe its constituents. The first key insight is that first two postulates already assume that, at least in some cases, the probability distribution of the
+	measurement outcomes of one system is independent of the other systems' (statistical independence).
+While we will see specifically how the Born implies this, it should be clear conceptually that we would not be able to talk about a system in the first place if we would not be able to characterize it independently.
+The second key insight is that, since all measurement
+	outcomes and all states of the subsystems plus the composite live in the same probability space, we will automatically have a map $M$ that for a pair of subsystems' states gives us the composite state that corresponds to the statistically independent case.
+These insights are enough to characterize mathematically the state space of the composite: the linearity given by the Hilbert space, together with the fact that the composite system is fully described by the observables of $A$ and $B$, allows to extend the construction to the composite states that do not satisfy statistical independence (the entangled states).
+So the work consists of two interrelated efforts: a physical argument that start from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.
+
+	We will mainly focus on
 	kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no superselection rules or
@@
 	since the tensor product formalism is very convenient and is often
-used also in these cases. Moreover, in accordance with the measurement
-postulate, we require that the probability distribution of the
-measurement outcomes of one system is independent of the other systems' (statistical
-independence). As detailed below, statistical independence is not an
-additional requirement: it is already contained in the measurement
-postulate.
+	used also in these cases, but it is something we need to bring out.
 
-From the above definition of composite system, it follows that there
-must be a map $M$ that connects the states of the subsystems to the
-states of the composite system. A quantum state is a ray in Hilbert
-space, namely a set of vectors. The map $M$ on states corresponds to a
-map $m$ on vectors. We then show that preparation independence and
+	While conceptually the derivation is straight forward, it requires the
+	use of projective space, which may be unfamiliar to the reader. This is
+	necessary because events in probability space do not correspond to vectors in the Hilbert space, but to subspaces.
+If $\psi$ is a vector, the state corresponds to the ray $\pj{\psi}$, which is the subspace spanned by $\psi$.
+Note that ``ray'' can be misleading since $\pj{\psi}$ is really a complex { \em plane }.
+In the same way, the event $X = x_0$ corresponds to the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$.
+The map $M$ derived from the probability space is a map between states as events in probability space, and therefore will be a map from rays to rays, which is a map between the projectives spaces.
+The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.
+
+	This is done by showing that preparation independence and
 	statistical independence imply three conditions on the map $m$:
@@
 
-We use the axiomatization of quantum mechanics based on the following
-postulates (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The state of a
+	We start with the axiomatization of quantum mechanics based on the following
+	postulates (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state of a
 	system is described by a ray $\pj{\psi}$ corresponding to a set of
@@
 	pairs) and returns a state in the projective space $\pj{\cal C}$ for
-the composite. The map $M$ that acts on rays corresponds to a map
+	the composite. To visualize this geometrically, given a ray (a complex plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex plane) in $\cal C$.  Our final goal will be to find a map
 	$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
-Hilbert spaces $\cal A$, $\cal B$ and $\cal C$.  Namely,
+	Hilbert spaces $\cal A$, $\cal B$ and $\cal C$ consistently with $M$. Namely,
 	$\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign indicates the
-elements in the projective space. We will prove that the map $m$ is
+	elements in the projective space.
+Again geometrically, $m$ takes a vector in each of $\cal A$ and $\cal B$, and returns a vector in $\cal C$ and we want this to be consistent with $M$ such that vectors picked from the same rays will return vectors in the same ray.
+We will prove that the map $m$ is
 	the tensor product. We focus on pure states here: the argument can be
@@
 	the joint probability of events $a$ and $b$ given $z$ is
-$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Then the events $a$ and $b$ are
-independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$ (this
-is the {\em definition} of independent events). Since the Born
-probability formula in postulate (b) contains only quantities of the
-system (the state and the observable's eigenstates {\em of the
-  system}), it implies the ``statistical independence'' between
+	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. This essentially the definition of conditional probability: $p(b|z\wedge a) = \frac{p(a\wedge b|z)}{p(a|z)}$. Two events $a$ and $b$ are defined to be 
+	independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$.
+However, it is more instructive to define independence with $p(b|z\wedge a)=p(b|z)$.
+This makes it more evident that the probability of $b$ does not depend on conditioning on $a$.
+If we imagine $a$ and $b$ to be outcomes for particular measurements, we are saying that the probability of observing $b$ is not a function of the value measured by $a$.
+We can recover the first definition from the second.
+In fact, $p(b|z\wedge a)=\frac{p(a\wedge b|z)}{p(a|z)} =p(b|z)$ which leads to $p(a\wedge b|z)=p(a|z)\:p(b|z)$.
+
+	We note that the Born
+	probability formula in postulate (b) is of the form $p(a|\psi_{\cal A}$ and contains only quantities of the
+	system (the pure state $\psi_{\cal A}$ and the observable's eigenstates $a$).
+Now, suppose that, unknown to us, an additional second system was prepared in a pure state $\psi_{\cal B}$ and consider $p(a|\psi_{\cal A}\wedge \psi_{\cal B})$.
+In what cases are allowed to use the Born rune?
+That is, in what cases we can say $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = |\<a|\psi_{\cal A}\>|^2$?
+Only in the cases where $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = p(a|\psi_{\cal A}$ are statistically independent.
+In this sense, postulate (b) already implies  ``statistical independence'' between
 	different systems if they are prepared independently. Namely, the
@@
 	the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
-colinear transformation by definition. In this case, the fundamental
+	colinear transformation by definition.
+Geometrically, recall that $M_b$ maps rays to rays.
+The fact that $M_b$ is colinear means that it also maps higher order subspaces to higher order subspaces (lines to lines, planes to planes, and so on) while preserving inclusion (if a line is within a plane, the mapped line will be within the mapped plane).
+In this case, the fundamental
 	theorem of projective geometry \cite{fun} applies, which tells us that
@@
 	\end{align}
-We can ignore the antilinear case \eqref{h2b}, which simply
-corresponds to the map $m$ mapping kets of one or both subsystems into
-bras before outputting the composite system. All these maps are
-trivially physically equivalent, since the dual-space description of a
-quantum system through bras is equivalent to the description using
-kets, or in other words, the wavefunction and its complex-conjugate
-contain identical information (for example, in his first papers
+	In this setting, the antilinear case \eqref{h2b} corresponds to a change of convention (much like a change of sign in the symplectic form for classical mechanics) and can be ignored.
+Given a Hilbert space, in fact, we can imagine replacing all wavefunctions and all the operators with their complex-conjugate.
+These changes would effectively cancel out leaving the physics being represented unchanged  (for example, in his first papers
 	Schr\"odinger used both signs in his equation: effectively writing
 	{\em two} equivalent equations with complex-conjugate solutions
-\cite{sch}). We can now repeat the same analysis for the second
+	\cite{sch}). Another way to see this is noting that, given any equation 
+	in terms of operators acting on ket vectors, one can write an
+	equivalent equation in terms of bra vectors by simply taking the
+	Hermitian conjugate of both sides of the first equation. Namely, the
+	equation A|psi>=B|phi> is equivalent to $<psi|A^dag=<phi|B^dag$.
+Mathematically, there is nothing preventing one to represent state preparation and kets to represent measurements.
+It is a matter of convention.
+We can now repeat the same analysis for the second
 	argument of $m$ to conclude that it is a bilinear map, condition (H2).
@@
 	where we used the orthonormality of the bases and the fact that
-$|a_i\otimes b_j\>$ is a basis of the tensor product space ${\cal
-  A}\otimes{\cal B}$. The function $\hat{m}$, then, is an isomorphism
-over all elements of the basis $|a_i\otimes b_j\>$. It is then an
-isomorphism between $\cal C$ and ${\cal A}\otimes{\cal B}$. Namely,
-$|m(a_i,b_j)\>_{\cal C} = |\hat{m}(a_i\otimes b_j)\>_{\cal C} \cong
-|a_i\otimes b_j\>_\otimes$, where the first two vectors (in $\cal C$)
-are isomorphic to the last vector (in the tensor product space ${\cal
-  A}\otimes{\cal B}$). Namely, $\hat{m}$ maps bijectively to the
-tensor product, which implies the isomorphism ${\cal C}\cong{\cal
-  A}\otimes{\cal B}$, i.e.~their equivalence through the map
-$\hat{m}$. Moreover, since ${\cal C}\cong{\cal A}\otimes{\cal B}$,
-then the map $m : \cal A \times \cal B \to \cal C$ is equivalent to
-the map $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$,
-i.e.~the maps $m$ and $\otimes$ are also equivalent through $\hat
-m$.$\square$
+	$|a_i\otimes b_j\>$ is a basis of the tensor product space ${\cal A}\otimes{\cal B}$.
+Since the function $\hat{m}$ is a linear function that maps an orthonormal basis of ${\cal A}\otimes{\cal B}$ to an orthonormal basis of $\cal C$, $\hat{m}$ is a an isomorphism (a bijection that preserves the mathematical structure) between ${\cal A}\otimes{\cal B}$ and $\cal C$.
+As ${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert spaces, they are mathematically equivalent and we can directly use the tensor product to represent the composite state space.
+This means that the map $m : \cal A \times \cal B \to \cal C$ is equivalent to the map $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$ in the sense that $m \circ \hat m^{-1} = \otimes$.$\square$
 

### 2020-11-07 Overviews/Classical mechanics/ClassicalMechanicsOverview.tex
Centering not needed
Fixing couple of equations

@@
 \urlstyle{same}
+\frenchspacing
 
@@
 \begin{figure}[h]
-	\centering
 	\includegraphics[width=\columnwidth]{Diagram.png}
@@
 
-We start with the simplest case, where one coordinate is sufficient to define the unit system, which means the following three conditions:
+We start with the simplest case, where one coordinate is sufficient to define the unit system, which means the following four conditions:
 \begin{enumerate}[noitemsep]
@@
 	\item we can arbitrarily change coordinate to $\hat{q}=\hat{q}(q)$
-	\item a change of coordinate induces a unique change over the remaining state variables $\hat{k}^j = \hat{k}^j(q, k^i)$.
+	\item a change of coordinate induces a unique change over the remaining state variables $\hat{k}^j = \hat{k}^j(q, k^i)$
+	\item the density is the same regardless of the coordinates used.
 \end{enumerate}
@@
 This means the volume form is proportional to $\omega^n$.
-This cannot be degenerate (i.e.~it must be nonzero for each infinitesimal volume), or some volumes would not contain states.
+This cannot be degenerate (i.e.~it must be nonzero for each infinitesimal volume), or some regions of the particle state space would not contain states, which is a contradiction.
 Therefore $\omega$ itself cannot be degenerate.
@@
 We first apply the assumptions to the motion of a single particle.
-That is, given a state $s_0$ at the initial time $t_0$, for each time $t$ there will be one and only one state $s_t$.
-There is one possible evolution $s(t)$ such that $s(t_0) = s_0$.
+Let $\lambda : \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
+Under the assumption, this will be uniquely identified by the initial state $s_0 = \lambda(t_0)$ at the initial time $t_0$.
 Secondly, we apply the assumption to the density in the sense that all the particles that start with the same initial state must end up in the same final state and vice-versa.
-That is, if $\rho(s(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(s(t), t) = \rho(s(t_0), t_0)$: the density must remain the same throughout the evolution.
+That is, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial particle at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$: the density must remain the same throughout the evolution.
 

### 2020-11-07 Overviews/Quantum mechanics/QuantumMechanicsOverview.tex

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-1}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\DeclareMathOperator{\tr}{tr}
\frenchspacing

\begin{document}

\title{Assumptions of Physics overview:  \\
	Quantum mechanics and irreducibility}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
We briefly show how quantum mechanics can be rederived and better understood as a consequence of irreducibility. The Schroedinger equation (i.e. unitary evolution) is recovered as deterministic and reversible motion while the measurement process is recovered as a special case of a process with equilibria (i.e. idempotent process, projection).  This work is an overview of some of the results of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
\end{abstract}

\maketitle

\section{Introduction}

The overall argument (see \cite{AoPPhy1,AoPBook} for more details) can be summed up in the following diagram that can be used as a guide throughout this note.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{Diagram.png}
\end{figure}

The assumptions lie on the left column. Each assumption leads to one or two key insights that progressively lead to the physical concepts in the middle column. Each of these is then mapped to its corresponding formal framework on the right. Note that ``quasi-static process'' and ``conserved density'' both independently lead to the same result of ``unitary evolution''.

\section{Irreducibility}

\textbf{Irreducibility assumption:} \emph{The state of the system is irreducible.} That is, giving the state of the whole system says nothing about the state of its parts.

Under this assumption the state of the system is automatically an ensemble over the state of the parts as preparation of the whole leaves the parts unspecified. For the same reason, the entropy of these ensembles must be the same, or some ensembles would provide more or less information about the parts. The whole task, then, is to characterize these ensembles without making specific assumptions on the parts.

Let $\mathcal{C}$ be the state space of the irreducible system. Let us call fragment a part of the irreducible system. The state of a fragment will be associated with a random variable uniformly distributed over the possible fragment states. As discussed in the context of classical mechanics(CITE), distributions over states must be invariant and symplectic manifolds are the only manifolds over which invariant distributions can be defined. As we cannot say anything about the state of the fragments, the dimensionality of this manifold must be irrelevant as long as it is even dimensional. For simplicity, we can choose a two-dimensional one. Therefore we are interested in the space of bi-dimensional uniform distributions formed by a pair of two random variables $A$ and $B$.

The values of the variables themselves are not relevant, as they are not physically accessible by assumption. However, the size of the system $\mu = \int \rho dA \wedge dB$ is relevant. Without loss of generality, we can rescale $A$ and $B$ such that the density $\rho$ is not only uniform but unitary: $\rho = 1$. This way the size of the system is directly proportional to the area covered by the random variables. In other words: the more fragments there are, the more each fragment can swap its state with another without changing the whole, the more uncertainty there is on the state of the fragment, the higher the variance of the random variables.

Since only linear transformations will preserve the uniform distribution, we look to those. These are translations, stretches and rotations. Translations do not lead to other physically distinguishable states since the exact values of $A$ and $B$ are not physically accessible. Stretching of the distribution will correspond to an increase of the size of the system, which is physically accessible. However, only the stretching of the area is of interest. So, without loss of generality, we can set $\sigma_A = \sigma_B = \sigma$ and we have $\mu \propto \sigma^2$. Rotations just change the correlations which, by themselves, are not physically accessible. However, under addition the correlations still result in differences in variance and, indirectly, the size of the system, and therefore are physically interesting. The space of transformations is therefore given by two parameters $a$ and $b$ such that:
\begin{equation}
\begin{aligned}
C &= a A + b B \\
D &= -b A + a B
\end{aligned}
\end{equation}
Equivalently, we can use the complex number $c = a + \imath b$ to characterize the transformation, which we can note as $\tau(c)$. The increase/decrease in size is given by $a^2 + b^2 = (a - \imath b) (a + \imath b) = c^* c$ and the change in correlation is given by the Pearson correlation coefficient $\rho_{A,\tau(c) A} = \cos \arg c$.

Putting it all together, we can characterize the state space $\mathcal{C}$ with a complex vector space. The linear combination represents the mixing of the different stochastic descriptions. Two vectors that only differ by a total phase are physically equivalent since a global change of correlation does not change the distribution.

We can define a scalar product $\langle \cdot | \cdot \rangle$ where the square norm induced corresponds to the size of the system (or equivalently to the strength of the random variable) and the phase difference corresponds to the correlations (the Pearson correlation coefficient). To see this, note the formal equivalence between the variance and norm rules under linear composition:
\begin{equation}
\begin{aligned}
\sigma^2_{X+Y} &= \sigma^2_{X} + \sigma^2_{Y} + 2 \, \sigma_{X} \sigma_{Y} \rho_{X,Y} \\
|\psi+\phi|^2&=|\psi|^2 + |\phi|^2 + 2 |\psi||\phi|\cos(\Delta \theta)
\end{aligned}
\end{equation}
The quadratic form, again, reflects the fact that the size of the system is proportional to the variance of a random variable. Since the size of the system is fixed, we use unitary vectors to represent actual states. The state of the system, then, is represented by a ray in a complex inner product space.

Lastly, we need to define an expectation operator that returns the average value for each physical quantity. This operator will have to be linear under linear combination of quantities:
\begin{equation}
	E[aX + bY | \psi] = aE[X | \psi] + bE[Y | \psi].
\end{equation}
 It will not be linear under linear combination of states:
 \begin{equation}
 	E[X | \psi + \phi] \neq E[X | \psi] + E[X | \phi].
 \end{equation}
Yet, it will have to be proportional to the increase in size and invariant under a total change in correlation: $E[X | \tau(c) \psi] = c^*c E[X | \psi]$. This leads us to associate to each physical quantity a linear Hermitian operator $X$ where $E[X | \psi] = \langle \psi | X | \psi \rangle$. An eigenstate $\psi_0$ of $X$ corresponds to a state where all the elements of the ensemble have exactly the same value. That is, $E[(X - \bar{x})^2 | \psi_0] = 0$.

Note that an inner product space can always be completed into a Hilbert space. This may, however, bring in objects that may not correspond to physical objects (i.e. infinite expectation for some quantities). In general, we believe it is better to regard the (possibly incomplete) inner product space as the physical state space and regard the completion as a mathematical device for calculation. For example, the Schwartz space seems more physically meaningful than the standard $L^2$ space as it gives finite expectation of all polynomials of position and momentum and, moreover, it is closed under Fourier transform.

\section{Process with equilibria}

\textbf{Process with equilibria assumption:} \emph{given an initial ensemble (i.e.~mixed state), the final ensemble is uniquely determined and remains the same if the process is applied again.}

Under this assumption, the process can be characterized by a projection operator. Let $\rho_1$ be the density matrix that characterizes a mixed state. Since the final mixed state must be uniquely determined by $\rho_1$, it will be $\mathcal{P}(\rho_1)$ for some operator $\mathcal{P}$. Similarly, if $\rho_2$ is another initial mixed state, its final operator will be $\mathcal{P}(\rho_2)$. Note that, given any observable $X$ the expectation $E[X|\rho_1] = \tr(X\rho_1)$ is the trace of $X\rho_1$. Similarly $E[X|\mathcal{P}(\rho_1)] = \tr(X \mathcal{P}(\rho_1))$.

We can always create statistical mixtures of the ensembles and we must have $E[X|a \rho_1 + b \rho_2] = a E[X|\rho_1] + b E[X|\rho_2]$ since these are classical mixtures. But since these are classical mixtures, the final state will also need to obey $E[X|a \mathcal{P}(\rho_1) + b \mathcal{P}(\rho_2)] = a E[X|\mathcal{P}(\rho_1)] + b E[X|\mathcal{P}(\rho_2)]$ for all possible $X$. Which means $\mathcal{P}(a \rho_1 + b \rho_2) = a \mathcal{P}(\rho_1) + b \mathcal{P}(\rho_2)$ Therefore the operator $\mathcal{P}$ is a linear operator. Moreover, the process applied twice must lead to the same result, which means $\mathcal{P}(\mathcal{P}(\rho)) = \mathcal{P}(\rho)$ for any $\rho$. That is, $\mathcal{P}^2 = \mathcal{P}$. Therefore $\mathcal{P}$ is a projection.

Suppose, now, that we want to measure a quantity $X$. We want the final outcome, the final ensemble, to be determined by the initial state, the initial ensemble. We also want the measurement to be consistent in the sense that, if it is repeated immediately after, it should yield the same result. Therefore the process will be a projection. We will also want that the process does not distort the quantity. That is, $E[X|\rho] = E[X|\mathcal{P}(\rho)]$. This means that the eigenstates of $X$ will correspond to equilibria of the process. Moreover, subsequent measurements must give the same value, not just the same mixture. That is, if $X_1$ is the random variable after the first instance of the process and $X_2$ is the random variable after the second instance, $P(X_2 = x| X_1 = x ) = 1$. This means that $E[(X - \bar{x})^2|\mathcal{P}(\rho)] = 0$ which means the eigenstates of $X$ are the only equilibria.

The measurement process is therefore simply a special case of a process with equilibria.

\section{Deterministic and reversible evolution}

\textbf{Deterministic and reversible evolution assumption:} \emph{given the present state of the system, all future (determinism) and past (reversibility) states are uniquely identified.}

Under this assumption, the process can be characterized by unitary evolution (i.e.~the Schrodinger equation). There are multiple different ways to see this. The first relates to the more general idea that all deterministic and reversible processes must be isomorphisms in the category of states. Since the state space is an inner product space, the isomorphism is unitary evolution.

The second, is that if there is a set of quantities $X_0$ at time $t_0$ that fully identify the state (i.e. the state is the only eigenstate of those quantities), then there must be a corresponding set of quantities $X_1$ that fully identify the state at time $t_1$. This means that the evolution maps basis to basis. Moreover, given the linearity of statistical mixtures, this will also mean that a statistical distribution over $X_0$ will have to map to the same distribution over $X_1$. Therefore the evolution must map linear combinations of that basis to the same linear combination. The evolution is a linear operator. Since the total size of the irreducible system cannot change, the operator must be unitary.

The third, is by constructing a quasi-static process from processes with equilibria, much like one does in thermodynamics.  The idea is that we have an infinitesimal time step, an initial state $\psi_t$ and a final state $\psi_{t+dt}$. We want $P(\psi_{t+dt} | \psi_t ) = 1$. This means that $|\langle \psi_{t+dt} | \psi_{t} \rangle|^2 = 1$. This can happen only if the difference between initial and final states is infinitesimal. That is, $\langle \psi_{t+dt} | \psi_{t} \rangle = 1 + \imath \epsilon dt$ where $\epsilon$ is a real number. Therefore, by convention, we can write $| \psi_{t+dt} \rangle = I + \frac{H dt}{\imath \hbar} | \psi_{t} \rangle$ where $H$ is a Hermitian operator.

Putting these perspectives together, time evolution is a unitary operator which can be written as $U=e^{\frac{H\Delta t}{\imath \hbar}}$. If we start in an eigenstate of $X$, that is $X | \psi_t \rangle = x_0 | \psi_t \rangle$ we will end in an eigenstate $\hat{X} | \psi_{t + \Delta t} \rangle = x_0 | \psi_{t + \Delta t} \rangle$ of another operator $\hat{X} = e^{\frac{H\Delta t}{\imath \hbar}} X e^{- \frac{H\Delta t}{\imath \hbar}}$.

In fact:
\begin{equation}
\begin{aligned}
e^{\frac{H\Delta t}{\imath \hbar}} X e^{- \frac{H\Delta t}{\imath \hbar}}| \psi_{t + \Delta t} \rangle
&= e^{\frac{H\Delta t}{\imath \hbar}} X e^{- \frac{H\Delta t}{\imath \hbar}} U | \psi_t \rangle \\
&= e^{\frac{H\Delta t}{\imath \hbar}} X e^{- \frac{H\Delta t}{\imath \hbar}} e^{\frac{H\Delta t}{\imath \hbar}} | \psi_t \rangle \\
&= e^{\frac{H\Delta t}{\imath \hbar}} X  | \psi_t \rangle \\
&= e^{\frac{H\Delta t}{\imath \hbar}} x_0  | \psi_t \rangle \\
&= x_0 U | \psi_t \rangle \\
&= x_0 | \psi_{t + \Delta t} \rangle
\end{aligned}
\end{equation}
This is consistent with assuming there is a quasi-static process that, at every $t$, has equilibria identified by $e^{\frac{H (t - t_0)}{\imath \hbar}} X e^{- \frac{H (t - t_0)}{\imath \hbar}}$. Note that, unlike thermodynamics, the equilibria during the evolution are not set by external constraints but by the system itself. That is, $X$ depends on the initial state of the system.

In this light, the measurement processes and the unitary processes can be seen as particular cases of the same type of processes, those with equilibria, which are defined as a black-box from initial to final state. This is consistent with the irreducibility assumption as the inability to describe the dynamics of the parts implicitly assumes that the dynamics of the parts is at equilibrium and sets a time-scale under which the further description of the system (i.e. non-equilibrium dynamics) would require describing the internal dynamics.

\bibliography{bibliography}

\end{document}

### 2020-11-09 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Edits and elaboration
derivation summary, derivation exposition edits, fundamentality, reorganization of sections

@@
 
-\usepackage{amsmath, hyperref, enumerate}
+\usepackage{amsmath, hyperref, enumerate, comment}
 \usepackage{amsfonts}
@@
 
-\item motivate everything upfront: going to show that Hamiltonian mechanics corresponds to two conditions being satisfied, while Lagrangian mechanics requires a three condition
+\item motivate everything upfront: going to show that Hamiltonian mechanics corresponds to two conditions being satisfied, while Lagrangian mechanics requires a third condition
 \item define the two assumptions, and give motivation
@@
 
-TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.
+% TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.
 
 Philosophers have recently shown that we can interpret Lagrangian and Hamiltonian mechanics as theoretically equivalent in a precise mathematical sense.
-Within a subclass of classical mechanical systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh2017}{Barrett2018}.
+Within a subclass of classical mechanical systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh}{Barrett2}.
 Categorical equivalence provides a natural interpretive criterion for theoretical equivalence because it yields mutual inter-translatability between theory formulations.
 At first glance, this equivalence seems to motivate the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain.
-This view would undermine prior debates that have privileged either Hamiltonian mechanics or Lagrangian mechanics on physical or metaphysical grounds \parencites[]{North2009}{Curiel2014}.
+This equivalence undermines prior positions that have privileged either Hamiltonian mechanics or Lagrangian mechanics on physical or metaphysical grounds \parencites[]{North}{Curiel}.
 Here, we will argue that there are indeed physical grounds for privileging Hamiltonian mechanics over Lagrangian mechanics, their categorical equivalence notwithstanding.
 Compared to Hamiltonian systems, an additional physical condition must be satisfied for a classical system to be Lagrangian.
-Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this conceptual difference has interpretive consequences even within the hyper-regular domain.
+Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this difference has interpretive consequences even within the hyper-regular domain.
 Furthermore, the construction at the heart of this argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density, along with a novel interpretation of the principle of stationary action within the Hamiltonian formalism.
@@
 
-\subsection{Define Classical Mechanics}
+\section{Defining `classical mechanical systems'}
 
-Talk about dissipative systems.
+Provide any relevant background on classical mechanics, specifically to clarify/define terminology.
+Might be able to just put this in the introduction proper, or the start of the next section.
+Make explicit how our terminology relates to other standard terminology in the philosophical literature.
+
+A classical Hamiltonian system is any system whose phase space is a symplectic manifold and that evolves according to Hamiltonian evolution (i.e. according to a symplectomorphism).
 
@@
 For in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $.
-This requires that the system is conservative [why exactly?], which is equivalent to the system being deterministic and reversible.
+This requires that the system is conservative, which is equivalent to the system being deterministic and reversible.
 Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
@@
 
-\subsection{Derivation}
+\section{Privileging Hamiltonian mechanics}
+
+Our framework develops a correspondence between three physical conditions and their associated mathematical formalisms.
+In this way, we straightforwardly interpret the physical significance of these mathematical formalisms.
+First, a system is \textit{infinitesimally reducible} provided that the states of its constituent, infinitesimal particles completely specify the overall state of the system.
+When this condition is satisfied, we can define a distribution $\rho$ over particle states.
+Furthermore, since particle states are independent of a choice of units, this distribution must be invariant.
+We show in Section~\ref{infinitesimal} that any infinitesimally reducible system has a phase space given by a symplectic manifold.
+The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
+
+Secondly, a system has \textit{deterministic and reversible evolution} provided that the present state of the system uniquely determines both its future and past states.
+This condition entails that the preceding invariant distribution is in fact a conserved phase space density.
+We can thus model such systems using Hamiltonian mechanics, defined on a phase space with a symplectomorphism.
+These two conditions thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution.
+Section~\ref{deterministic} provides this derivation in detail.
+
+Finally, a system satisfies a further condition---motion equivalence---when its trajectory in physical space determines and is determined by its evolution in phase space.
+This condition leads to a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics in the hyper-regular domain (Section~\ref{motion}).
+Hence, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
+To see this, note that some systems fail to satisfy the condition of motion equivalence, i.e.
+it is not always true that we can determine the evolution of a system in phase space from its trajectory in physical space.
+For instance, knowledge of the trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from its trajectory alone.
+Hence, the set of classical systems that satisfy this further physical condition is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics.
+On these grounds, we will argue in Section~\ref{fundamentality} that Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
+
+% Overall summary: A classical system is an infinitesimally reducible system.
+The state space of these systems can be modeled by a symplectic manifold (the physical assumption of infinitesimal reducibility corresponds to the mathematical structure of symplectic manifolds).
+Classical Hamiltonian systems satisfy the additional conditions of having deterministic and reversible evolutions.
+These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it a contact manifold.
+Classical Lagrangian systems satisfy a further assumption: keeping track of the trajectories in physical space is equivalent to tracking the evolution of states in phase space.
+
+Figure~\ref{diagram} summarizes our characterization of Hamiltonian and Lagrangian mechanics, in terms of physical conditions and corresponding mathematical structures.
+The left column lists the three physical conditions: infinitesimal reducibility, determinism/reversibility, and motion equivalence.
+Each condition entails physically significant properties, described in the middle column.
+Each of these physical properties corresponds to a formal mathematical framework, described in the right column.
+Section~\ref{derivation} justifies the logical connection between these physical conditions and corresponding mathematics.
+Before delving into this derivation, Section~\ref{fundamentality} discusses its philosophical significance.
 
@@
 	\includegraphics[width=\textwidth]{Diagram.png}
+\caption{Physical conditions and corresponding properties and formalisms}
+\label{diagram}
 \end{figure}
 
-Overall summary: A classical system is an infinitesimally reducible system.
-The state space of these systems can be modeled by a symplectic manifold (the physical assumption of infinitesimal reducibility corresponds to the mathematical structure of symplectic manifolds).
-Classical Hamiltonian systems satisfy the additional conditions of having deterministic and reversible evolutions.
-These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it a contact manifold.
-Classical Lagrangian systems satisfy a further assumption: keeping track of the trajectories in physical space is equivalent to tracking the evolution of states in phase space.
-Note that it is not always true that we can determine the evolution in phase space from the trajectory in physical space.
-For instance, knowledge of the trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from its trajectory alone.
-Hence, the set of classical systems that satisfy this further physical condition is a proper subset of the classical systems that satisfy the assumptions for Hamiltonian mechanics.
+\subsection{Fundamentality and structure}
+\label{fundamentality}
 
-\subsubsection{Infinitesimal reducibility}
+In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by ``fundamental." Of the various notions of fundamentality, there is at least the following non-mysterious definition: a formalism is more fundamental provided that it is logically more general.
+In this case, the Lagrangian framework is less general because it requires that an additional condition hold, compared to the Hamiltonian framework.
+Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.
+In particular, the assumption of motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.
+This notion of fundamentality leads as well to a concomitant notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied.
+In Section~\ref{distribution}, we will discuss another sense in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.
+% Way to motivate that the motion equivalence assumption further constrains the Hamiltonian: the Hamiltonian already defines a relationship between conjugate momentum and velocity (through ($dq/dt = \partial H / \partial p$) ).
+Motion equivalence provides a further constraint on this relationship, and hence a further constraint on the Hamiltonian.
+\\
+
+% As defined, our notions of fundamentality and structure are not metaphysical in nature.
+
+[Explain how our notion of comparative structure avoids the objections of Halvorson/Swanson against North's comparative notion of amount of structure.] \\
+
+We do not endorse a metaphysical notion of fundamentality, on which a formalism is more fundamental provided that it is more joint-carving.
+For instance, North provides the following definition of structure: ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" \parencites*[66]{North}.
+While Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{distribution}), we will also argue that classical mechanics deploys physically significant representations that are not intrinsic to mechanical systems.
+Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.
+Whether Hamiltonian phase space structure is metaphysically fundamental does not matter for the questions we seek to answer.
+Regardless of its metaphysical status, Hamiltonian mechanics is epistemically, conceptually, and physically more fundamental.
+
+% Part of the following paragraph seems redundant given the preceding. Think about how to combine.
+Methodologically, we agree with North that it is important to identify invariants and more generally ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}.
+Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North}.
+However, whereas North proceeds mainly from the mathematical structure to claims about physical significance, we seek to interpret the mathematics by beginning with physically significant assumptions.
+We seek to determine which physical assumptions are necessary and sufficient for supporting a corresponding mathematical structure.
+This does not entail that additional mathematical structure is necessarily physically insignificant or surplus structure, just that it may correspond to an additional physically significant assumption that might apply to the system under consideration.
+As we will see, this characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system.
+Hence, whereas North views the Lagrangian structure as surplus or redundant (``excess, superfluous structure"), we take the Lagrangian structure as unnecessary but still physically significant, corresponding to an additional physical assumption that applies to some classical systems.
+
+Furthermore, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references].
+Each of these frameworks has different pragmatic costs and benefits.
+Hence, unlike North, we are comfortable acknowledging that the mathematical structure that you ``need" might not be unique.
+There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions.
+Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others.
+And in some sense, each of these formalisms references some ``unphysical" coordinates or representational means.
+Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics.
+In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.
+% One con of this extended phase space formulation: the Hamiltonian constraint is both a constraint and the generator of time evolution; hence, it must be defined in the neighborhood of the actual states, so the values it provides outside of these neighborhoods should be unphysical, yet they encode the equations of motion.
+% Should we remark on the claim that the only structure we need is whatever is sufficient to define densities locally? Hence, we need just the symplectic form but not the cotangent bundle?
+% E.g.
+in the extended phase space formulation, the equations of motion come from derivatives of the Hamiltonian constraint over the variables, but this constraint is defined on unphysical points.
+These unphysical points are plausibly a mathematical artifact, although they supply a nice tool/method for understanding physically significant aspects of the representation.
+
+Claim to think about: GC was arguing on 10-12 that you could handle classical systems in the Hamiltonian framework with just a differential manifold and no symplectic structure.
+This would contradict North's indispensability claim that Hamiltonian symplectic structure is necessary for treating classical systems.
+E.g., North footnote 31: ``The symplectic form is crucial to determining the motion, since the same Hamiltonian can induce different flows for different symplectic forms" [76].
+North takes this as an argument for why we cannot eliminate the symplectic form in an attempt to get at even less structure.
+We need it to define the dynamics.
+
+% North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory.
+Structure comprises those quantities that remain intact when we transform one allowable coordinatization of the laws into another" [78]
+
+% North definition of structure: ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" [66].
+So presumably, North would view the invariant phase space density as part of this fundamental structure.
+
+% North's epistemic method for figuring out the structure: ``look at what is invariant under allowable transformations"; this tells us what the structure is [73].
+
+\section{Derivation}
+\label{derivation}
+
+Sections~\ref{infinitesimal}-\ref{deterministic} show how to derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution.
+Section~\ref{motion} shows how systems that satisfy the further condition of motion equivalence can be treated within the formalism of Lagrangian mechanics.
+
+\subsection{Infinitesimal reducibility}
+\label{infinitesimal}
 
@@
 \begin{quotation}
-Infinitesimal reducibility assumption: the state of the system is reducible to the state of its infinitesimal parts.
+\noindent
+\textit{Infinitesimal reducibility}: the state of the system is reducible to the state of its infinitesimal parts.
 That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the states of those parts, and so on.
@@
 A classical system then consists of a collection of classical particles.
-Let $\mathcal{S}$ be the state space for the particles [all of the particles constituting the whole system?
-or for each particle individually?].
+Let $\mathcal{S}$ be the state space for the particles, so that each point in $\mathcal{S}$ describes a possible state of a classical particle.
 We assume that this state space $\mathcal{S}$ is a manifold.
@@
 That is, the state of the whole system is a distribution over its infinitesimal parts.
-If $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space $\mathcal{S}$.
+This distribution describes the proportion of classical particles in any particular state.
+Formally, if $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space $\mathcal{S}$.
 %[is region U in the state space C or S. I assume in S.].
 
-On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function, i.e.
-it must depend on the point [the point $c $?
-] and not on the coordinates.
+We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state (in manifold terminology, these are known as coordinates).
+Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$, yielding a real-valued vector for any given particle state, with $n$-many degrees of freedom.
+We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
+For instance, $q$ might be position, defining units of meters.
+On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function.
 This is because the distribution is a function only of the states, rather than of the particular choice of state variables, i.e.
 coordinates.
-[Explain why these two physical conditions hold for the distribution.
-Also, explain what it means to transform as a density.] We will now show that classical phase space (i.e.
+To transform as a density means that $\rho_c$ changes as a function of the Jacobian determinant that relates the new coordinates to the old coordinates.
+To transform as a scalar means that the distribution $\rho_c$ is invariant under all coordinate transformations, so that the Jacobian determinant must in fact equal one.
+Hence, $\rho_c$ is a function only of the point $c$ and does not depend on the choice of coordinates.
+We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
+In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a density and scalar function.
+% Classical phase space (i.e.
 a symplectic manifold) is the only space that satisfies these constraints on the distribution.
 
-We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state (in manifold terminology, these are known as coordinates).
-We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
-For instance, $q$ might be position, defining units of meters.
-Starting with the simplest case, suppose that one coordinate suffices to define the unit system for the particle state space.
+The argument begins with the simplest case: suppose that a single coordinate $q $ suffices to define the unit system for the particle state space.
 Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number.
 We will show that, in fact, there can only be a single $k^i$ coordinate in this case, conjugate to the single $q$ coordinate.
-Our derivation assumes---on physical grounds---that we can (i) arbitrarily change the unit defined by $q$ without changing the density $\rho_c$ and that (ii) $q$ fully defines the units for all state variables.
-%[Give a concrete example of a state variable defining a unit, along with physical motivation for why state variables play the role of defining units (since we are treating this as a physically grounded assumption).]
-% How many $k^i$ can we have?
+Assuming that the particle density $\rho_c$ is a scalar, it follows that we can arbitrarily change the unit defined by $q$ without changing the density $\rho_c$.
+% We further assume that $q$ fully defines the units for all state variables.
+[Is there a way to physically motivate this assumption, beyond this being a constitutive role of the position-like state variables?]
 
-[We need to introduce the definition of state variable evolution $s$, since we appeal to it below.
-I take it that the definition we are working with is that future states are labeled by the original state variable values (i.e.
-transported state variables)?
-Section IV.C of long paper.]
+[We might be able to shorten part of the following argument simply by noting that the Jacobian determinant must equal unity, as explained verbally in a prior paragraph.
+The preceding verbal justification might be sufficient, i.e.
+that the distribution transforms as a density and a scalar.
+We could then focus on the trickier part of showing that there is only a single $k^i$ variable, based on the form of the Jacobian determinant.]
 
-Now, suppose that we change the units by transforming the coordinate $q$ to a new coordinate $\hat{q}$ that is a function of the old coordinate: $\hat{q}=\hat{q}(q)$.
+Now, suppose that we change the units by transforming the coordinate $q$ to a new coordinate $\hat{q}$.
+Since---by assumption---$q$ fully defines the unit system, $\hat{q}$ must do so as well, making it a function just of the old coordinate: $\hat{q}=\hat{q}(q)$.
 Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
-We have $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$.
-Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must be equal to 1.
-Suppose there is only one state variable, so that there are no $k^i$ variables.
+As defined, $\rho_c$ is a function from the particle state space $\mathcal{S}$ to the reals.
+We can define $\rho_c$ on the units by taking advantage of an invertible function $s: \mathbb{R}^n \to \mathcal{S}$ mapping a particular value of the units $\hat{\xi}^b$ to a point in the particle phase space $\mathcal{S}$.
+Hence, by definition $\rho_c(\hat{\xi}^b):= \rho_c(s(\hat{\xi}^b))$.
+[Is this function $s$ the inverse of the coordinate mapping $\xi^b$?].
+Then, since the physical state is independent of the coordinates, the state $s(\hat{\xi}^b)$ is the same state as $s(\xi^a) $ defined in terms of the original units $\xi^a$.
+Again, by definition, $\rho_c(s(\xi^a))=\rho_c(\xi^a) $.
+Since the distribution transforms as a density, it follows that $\rho_c(\xi^a)$ must equal the product of the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| $ times the density defined over the transformed units: $\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$.
+We can summarize the argument so far as a chain of equalities:
+
+\begin{equation*}
+\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)
+\end{equation*}
+
+\noindent
+Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one.
+Assume for contradiction that there is only one state variable, so that there are no $k^i$ variables.
 Then we would have $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right| = 1$.
 But this would mean the unit change cannot be arbitrary.
-Therefore we must have two or more variables, so that there is at least one $k^i$ variable.
-The requirement that the Jacobian determinant equals one places a constraint on how the state variables $k^i$ must change to compensate the change from $q$ to $\hat q$.
-Note that, since $\hat{q}$ depends only on $q$, the Jacobian is an upper triangular block matrix, so $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$ [it might be worthwhile to spell this out in more detail].
-For the Jacobian determinant to equal one, we must therefore have $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$.
-This places a constraint only on the determinant of the transformation.
-Suppose there are three or more variables.
-This constraint would then be insufficient to recover the transformation uniquely, and therefore $\hat q$  would not fully define the units for all other state variables (contrary to assumption).
-This means there must be exactly two state variables: $q$ and a single $k$.
-So in this case, coordinate-independent areas and densities can only be defined on a two-dimensional manifold.
+Therefore we must have two or more state variables, so that there must be at least one $k^i$ variable.
+
+Next, we show that in fact there must be exactly one $k$ variable, conjugate to the single $q$ variable.
+Using the fact that the Jacobian determinant equals one, we determine how the state variables $k^i$ must change in order to compensate the change from $q$ to $\hat q$.
+Since $\hat{q}$ depends only on $q$, $\frac{\partial \hat{q}}{\partial k^a} = 0$.
+Hence, the Jacobian is an upper triangular block matrix:
+
+\begin{equation}
+%\begin{aligned}
+|J| = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \begin{vmatrix}
+\frac{\partial \hat{q}}{\partial q} & \frac{\partial \hat{q}}{\partial k^a} \\
+\frac{\partial \hat{k}^b}{\partial q} & \frac{\partial \hat{k}^b}{\partial  k^a} 
+\end{vmatrix} = \begin{vmatrix}
+\frac{\partial \hat{q}}{\partial q} & 0 \\
+\frac{\partial \hat{k}^b}{\partial q} & \frac{\partial \hat{k}^b}{\partial  k^a}
+\end{vmatrix}
+%\end{aligned}
+\end{equation}
+
+\noindent
+The determinant of a triangular block matrix equals the product of determinants of the block matrices along the diagonal: $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| =  \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right|$.
+For the Jacobian determinant to equal one, we must therefore have $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| =   \left|\frac{\partial \hat{q}}{\partial q}\right|^{-1}$.
+Since $\hat q$ is a function only of $q$, $\frac{\partial \hat{q}}{\partial q}$ is a total derivative, so we obtain the inverse by swapping the numerator and denominator.
+Hence, we see that $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$.
+%  $\left|\frac{\partial \hat{q}}{\partial q}\right|^{-1} = \left|\frac{\partial q}{\partial \hat q} \right|$.
+
+This equation provides only a single constraint, placed on the determinant of the transformation from $k^a$ to $\hat{k}^b$.
+Hence, if there were three or more $\xi^a$ variables, this constraint would be insufficient to recover the transformation of the $k^a$ uniquely.
+Yet in this case, $\hat q$  would not fully define the units for all other state variables, which is contrary to assumption.
+Consequently, there must be exactly two state variables: $q$ and a single $k$.
+So in this case of a single $q$ variable, coordinate-independent areas and densities can only be defined on a two-dimensional manifold.
+
+\begin{comment}
+\begin{equation}
+\left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right| - \left|0\right| \left|\frac{\partial \hat{k}^b}{\partial q}\right|
+= \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right|.
+\end{equation}
+ so $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$: 
+\end{comment}
 
@@
 
-We can now show that $\mathcal{S}$ is a symplectic manifold.
+We can now show that $\mathcal{S}$ must have at least the structure of a symplectic manifold.
 Invariant densities can only be assigned to infinitesimal areas.
-Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$ [clarify which phase space $\Sigma$ belongs to].
+Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$, for $\Sigma \subset \mathcal{S}$.
 Because degrees of freedom are---at least locally---independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
 This means the volume form is given by $\omega^n$.
-This cannot be degenerate, or some volumes would not contain states [motivate this claim or].
+Importantly, the volume form cannot be degenerate because otherwise there would be a volume in state space that could not contain states.
+But such a region of state space would---by definition---not be a volume.
 Therefore $\omega$ itself cannot be degenerate.
 As the degrees of freedom are independent, the number of states on a surface does not change if we translate it across independent degrees of freedom.
-If we imagine a parallelepiped, the integral over the surface must be zero (the integrals over opposite sides are equal and opposite).
+If we imagine a parallelepiped, the integral over the surface must be zero (the integrals over opposite sides are equal and opposite) [explain/justify this more clearly.
+Is this coming from the two-form being antisymmetric?].
 Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold.
@@
 
-The conclusion is that, under the infinitesimal reducibility assumption, the state space of the system is a coordinate-invariant distribution over the state space of the infinitesimal parts [what does it mean for the state space $\mathcal{C} $ to itself be a distribution ?].
-If the state space of the particles is a manifold, then it is a symplectic manifold, which is the only type of manifold that supports coordinate-invariant distributions [do we have a reference to substantiate the ``only" claim?
-Has the proof above really shown that this is the only kind of manifold that would support a coordinate-invariant distribution].
+The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is a coordinate-invariant distribution over the state space $\mathcal{S}$ of its infinitesimal parts.
+If the state space of the particles is a manifold, then it must be a symplectic manifold: symplectic manifold structure is necessary for the distribution to be coordinate-invariant.
+% Which is the only type of manifold that supports coordinate-invariant distributions
 
-\subsubsection{Deterministic and reversible evolution}
+\subsection{Deterministic and reversible evolution}
+\label{deterministic}
 
@@
 \begin{quotation}
-
-Deterministic and reversible evolution assumption: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.
+\noindent
+\textit{Deterministic and reversible evolution assumption}: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.
 
@@
 
-[Should we footnote that these two assumptions leave out some classical systems, e.g. Norton's dome, Earman's space invaders?]
+[Should we footnote that this condition leaves out some classical systems, e.g. Norton's dome, Earman's space invaders?]
 
 In our case, this applies to both the state of the whole system and to the state of its parts.
-Let $\lambda : \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
-Under the assumption, this will be uniquely identified by the initial state $s_0 = \lambda(t_0)$ at the initial time $t_0$.
-[is $s$ an evolution that transports the state variables, i.e.
-labeling future states by the original state variables so that the initial time is transported to itself?
-I think we need to explain more about how evolutions are defined].
-Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial particle [what exactly does `initial particle' mean?
-Do we mean the initial state of the composite system or just a particular particle?] at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e.
+Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
+Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$, at the initial time $t_0$, uniquely identifies the state of the system.
+Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial state at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e.
 the density remains the same throughout the evolution.
-In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa [could two different particles occupy the same initial state?].
+In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa.
+%[could two different particles occupy the same initial state?].
+%[what exactly does `initial particle' mean? Do we mean the initial state of the composite system or just a particular particle?]
 
@@
 That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (otherwise, volumes would not be mapped to equal volumes, which would violate determinism).
-The evolution is a symplectomorphism and corresponds to Hamiltonian evolution.
+The evolution is therefore a symplectomorphism and corresponds to Hamiltonian evolution.
 Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
+In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.
 
-\subsubsection{Motion equivalence}
+\subsection{Motion equivalence}
+\label{motion}
 
-Having recovered classical Hamiltonian mechanics, the laws of evolution become further restricted when the following condition holds:
+Having recovered classical Hamiltonian mechanics, the laws of evolution are further restricted when the following condition holds:
 
-Motion equivalence assumption: the motion of the system (i.e. trajectories in physical space-time) is enough to recover its dynamics (i.e. evolutions in state space) and vice-versa.
+\begin{quotation}
+\noindent
+\textit{Motion equivalence assumption}: the motion of the system (i.e. trajectories in physical space-time) suffices to recover its dynamics (i.e. evolutions in state space) and vice-versa.
+\end{quotation}
 
@@
 a state variable that defines a unit.
-In fact, the trajectories can be fully described by these units and only these units.
-So we can say that $q^i=x^i$, each $q^i$ will be paired with a conjugate $p_i$ and each state $\{q^i, p_i\}$ will be mapped to one and only one trajectory.
-At each point $x^i$, then, infinitely many trajectories must pass, one for each combination of $\{p_i\}$.
-Since the trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
+In fact, the trajectories can be fully described by these units and only these units [Say more to motivate this claim.
+Epistemic claim?
+Claim about measurement?].
+It follows that $x^i = q^i$, each $q^i$ is be paired with a conjugate $p_i$, and each state $\{q^i, p_i\}$ corresponds to one and only one trajectory in physical space.
+At each point $x^i$, then, there must be infinitely many trajectories, one for each possible combination of $\{p_i\}$.
+Since these trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
 If the equations of motions were such that $v^i=v^i(q^i)$, then motion equivalence would fail as the full trajectory would be determined only by the $q^i$.
-So we must have $v^i=v^i(q^i, p_i)$.
-This relationship must be invertible or motion equivalence would fail.
-At any given time, then, we must have the following relationships:
+So we must have $v^i=v^i(q^j, p_k)$.
+This relationship must be invertible or motion equivalence would fail [justify why: need to determine motion in physical space from phase space evolution].
+For systems that satisfy the motion equivalence condition, then, the following relationships hold at any given time:
 \begin{equation}
@@
 
-In this case, which we call \textit{weak equivalence}, $v^i(q^j, p_k)$ must be invertible at every $q^i$.
-Therefore, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
-Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ must be non-zero everywhere, and therefore must have the same sign which we take to be positive.
+In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^i$.
+Therefore, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible [Shouldn't the denominator of the Jacobian really be $\xi_j$, since it is with respect to both the q's and the p's?.
+We don't seem to justify that the Jacobian equals the Hessian until after calculating the determinant below].
+Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ equals the Jacobian.
+Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere, which we take to be positive.
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).
@@
 \end{vmatrix} \\
-&= \left|\delta^i_j\right| \left|\frac{\partial v^i}{\partial p_j}\right| - \left|0\right| \left|\frac{\partial v^i}{\partial q^j}\right|
+&= \left|\delta^i_j\right| \left|\frac{\partial v^i}{\partial p_j}\right| 
 = \left|\frac{\partial v^i}{\partial p_j}\right|.
@@
 Similar considerations will also hold for marginal distributions (i.e.
-distributions on a subset of the coordinates) which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
+distributions on a subset of the coordinates), which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
 We set:
@@
 \end{equation}
-where $m$ is the unit conversion constant between velocity and conjugate momentum while $g_{ij}$ represents the linear dependency.
+where $m$ is the unit conversion constant between velocity and conjugate momentum while $g_{ij}$ represents the linear dependency [of what on what?].
 
@@
 
-\section{Argument for privileging Hamiltonian mechanics}
-
-\subsection{Relevant formal background from assumptions of physics framework}
-
-%On the notion of more or less structure, treated in such a way that we avoid the objections of Halvorson/Swanson against North's notion of amount of structure:
-Our comparisons of structure arise from logical chains of strict entailment, wherein something has more structure if and only if it is less general, i.e.
-if and only if it is a specialization of a more general case.
-So in this case, the Lagrangian framework has more structure than the Hamiltonian framework because Lagrangian mechanics requires an additional assumption.
-This difference in the number of assumptions corresponds to a difference in the number of commitments.
-When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?).
-In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.
-
-Hamiltonian mechanics makes fewer assumptions on the physical system being studied than Lagrangian mechanics.
-Lagrangian is a subset than Hamiltonian.
-Makes ``amount of structure" more rigorous.
-Hamiltonian defines the relationship between conjugate momentum and velocity ($dq/dt = \partial H / \partial p$), so kinematic assumption constrains the Hamiltonian.
-
-Claim to think about: GC was arguing on 10-12 that you could handle classical systems in the Hamiltonian framework with just a differential manifold and no symplectic structure.
-This would contradict North's indispensability claim that Hamiltonian symplectic structure is necessary for treating classical systems.
-E.g., North footnote 31: ``The symplectic form is crucial to determining the motion, since the same Hamiltonian can induce different flows for different symplectic forms" [76].
-North takes this as an argument for why we cannot eliminate the symplectic form in an attempt to get at even less structure.
-We need it to define the dynamics.
-
-\subsection{The significance of invariant distributions}
+\section{The physical significance of invariant distributions}
+\label{distribution}
 
@@
 
-Hamiltonian mechanics is ``fundamental" because allows to show that densities are coordinate independent quantities and therefore physical quantities.
+Hamiltonian mechanics is ``fundamental" because it shows that densities are coordinate independent quantities and therefore physical quantities.
 Canonical pairs (position and conjugate momentum) allows a simpler expression/comparison of densities over space.
@@
 
-Can characterize this argument as physically or epistemically privileging Hamiltonian mechanics
+While it is important to provide a physical interpretation of invariants, this is not to say that only invariant quantities are physically meaningful.
+For instance, the coordinates q define the units, and units are physically significant despite not being invariant.
+Units are physically significant because they allow us to assign numbers to a physical quantity.
+They are physically significant in part because they correspond to measurable quantities.
+Whereas North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}, some choices of the coordinates q are less physically meaningful because they are not properly connected with measurement.
+Some arbitrary mixings of q's and p's preserve the symplectic manifold but are not as physically significant.
+For this reason, the q's are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative p's are not.
+Whereas symplectic geometry creates a mathematical, syntactic symmetry between the q and p variables, this is not a physically significant symmetry from the standpoint of measurement.
+
+Moreover: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
+(In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).
+% It follows from the way in which position quantities fix the units that $\dot{x} = v $.
+It is built into the notion of a dynamical system that half of the state variables are derivatives of the other half.
+
+\section{Interpreting Lagrangian mechanics}
+
+The action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
+It also holds when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
+Also shows that Lagrangian is ``unphysical" and depends on gauge.
+
+Upshot: can better understand Lagrangian mechanics as a special case of Hamiltonian mechanics.
+Explain how the geometrical understanding of the stationary action in extended phase space should meet any criterion of ``naturalness" or ``natural construction." Lagrangian mechanics is actually less natural because contingent on having a bijection between velocity and conjugate momentum.
+
+%Discussion of gauge transformations in the context of Lagrangian vs.
+Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):
+
+Claim to clarify: motion equivalence allows us to understand the gauge redundancy of the Lagrangian.
+However, it is not true that the Lagrangian has these redundancies in virtue of the system satisfying motion equivalence.
+
+In some sense (to be made more precise), gauge transformations are more fundamental in Lagrangian mechanics than Hamiltonian mechanics.
+The conjugate momentum $p $ functions as a dummy variable in virtue of being a gauge quantity.
+From motion equivalence, the Lagrangian has more invariant structure than the Hamiltonian [not sure what this really means].
+The Lagrangian has the advantage of being invariant under coordinate transformations, whereas the Hamiltonian is not a coordinate invariant.
+Instead, the Hamiltonian varies as the time component of a 4-co-vector.
+Nevertheless, the Lagrangian itself is physically problematic because it is a gauge-dependent quantity.
+The p variables are not of fundamental significance because we can perform a gauge transformation on the p's such that the q and $\dot{q} $ variables remain the same.
+For systems where motion equivalence holds, we are really measuring the q and $\dot{q} $ variables, indicating that this is a genuine gauge transformation of the momentum variables.
+Hence, privileging Hamiltonian mechanics does not entail, pace North, that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
+Instead, position and kinetic momentum remain physically privileged, with conjugate momentum relegated to a gauge quantity (despite its role in the Hamiltonian formalism).
+% But there is a sense in which on the extended phase space, the Hamiltonian constraint is an invariant, yielding a coordinate-invariant way of handling systems with Hamiltonian mechanics. (?) 
+%Since the q and $\dot{q} $ variables fix the physical state, this indicates that transformed conjugate momentum is a gauge quantity. 
+%North argues that ``momentum is a truly fundamental quantity, on a par with position" [77].
+
+Issue of the mathematical structure that is necessary/sufficient for Lagrangian mechanics:
+North footnote 20 and appendix: it is not generically true that the state space of Lagrangian mechanics comes equipped with a metric.
+Footnote 22: ``a more general, coordinate-free version" of Lagrangian mechanics ``does not make explicit use of this metric"; discussed in appendix A.
+North: ``the square of the Riemannian line element is what gives the allowable sets of coordinate transformations, the ones that preserve the Lagrangian equations of motion.
+A volume element would not suffice for this" [75].
+How does this connect with motion equivalence and its associated mathematical structure?
+
+North assumes that formulating a Lagrangian theory requires specifying a Riemannian metric; according to Curiel, this is neither necessary nor sufficient.
+Furthermore, the physical significance of a Riemannian metric in the context of a Lagrangian theory is unclear: two different Lagrangian's can have the same Lagrangian vector field and yet induce inequivalent Riemannian metrics
+
+This difference in the number of assumptions corresponds to a difference in the number of commitments.
+When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?).
+In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.
+
+%Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics.
+Curiel frequently appeals to naturalness and a distinction between natural vs.
+ad hoc/artificial constructions [figure out what his definition of ``naturalness" is, if he provides one].
 
@@
 
-Part of our response: The action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
-It also holds when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
-Also shows that Lagrangian is ``unphysical" and depends on gauge.
-
-Upshot: can better understand Lagrangian mechanics as a special case of Hamiltonian mechanics.
-Explain how the geometrical understanding of the stationary action in extended phase space should meet any criterion of ``naturalness" or ``natural construction." Lagrangian mechanics is actually less natural because contingent on having a bijection between velocity and conjugate momentum.
-
-On its own, the kinematic/motion of a system in physical space is not sufficient for defining the system's dynamics.
+On its own, the motion of a system in physical space is not sufficient for defining the system's dynamics.
 The kinematics/motion does not suffice to determine the forces acting on the system.
 For it is essential to be able to distinguish the forces acting on an inertial system from fictitious forces arising in the description of a non-inertial system.
-Hence, unless one knows that the kinematic equivalence principle is satisfied, Lagrangian mechanics faces an underdetermination problem: it is unable to distinguish between certain inertial vs.
+Hence, unless one knows that the motion equivalence principle is satisfied, Lagrangian mechanics faces an underdetermination problem: it is unable to distinguish between certain inertial vs.
 non-inertial systems that are physically distinct.
@@
 
-\subsection{Discussion of relation to categorical equivalence proofs}
+\section{Discussion of relation to categorical equivalence proofs}
 
@@
 
-\section{A non-metaphysical notion of fundamentality}
-
-%Different notions of fundamentality, perfectly natural properties, naturalness, etc.
-
-%Joint carving, perfectly natural properties.
-Ontological notion of fundamental.
-Clearly state why this is not ``good" for this context.
-We are agnostic.
-One may use following points to argue whatever they want.
-
-Whereas North \parencites*[76]{North2009} seeks to determine the `real' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.
-Whether Hamiltonian phase space structure is metaphysically fundamental is neither here nor there for the questions we seek to answer.
-For regardless of its metaphysical status, Hamiltonian mechanics is epistemically, conceptually, and physically more fundamental.
-% Similarly, Curiel advocates neutrality on these metaphysical issues when it comes to the conceptual analysis of a physical theory's structure (304).
-% On the basis of the fundamentality of Hamiltonian mechanics, North argues that ``we should infer that symplectic structure is the `real' state space structure of the theory.
-We should infer that this structure corresponds to the `real' physical structure of a classical mechanical world" [76]
-%``there is an objective fact about what structure exists, there is a privileged carving of nature at its joints: along the lines of its fundamental physical structure" [82].
-
-Methodologically, we agree with North that it is important to identify invariants and more generally ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North2009}.
-Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North2009}.
-However, whereas North proceeds mainly from the mathematical structure to claims about physical significance, we seek to interpret the mathematics by beginning with physically significant assumptions.
-We seek to determine which physical assumptions are necessary and sufficient for supporting a corresponding mathematical structure.
-This does not entail that additional mathematical structure is necessarily physically insignificant or surplus structure, just that it may correspond to an additional physically significant assumption that might apply to the system under consideration.
-As we will see, this characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system.
-Hence, whereas North views the Lagrangian structure as surplus or redundant (``excess, superfluous structure"), we take the Lagrangian structure as unnecessary but still physically significant, corresponding to an additional physical assumption that applies to some classical systems.
-% % Similarly, we agree with the general methodology of Curiel as well: ``One task of the philosopher of physics is to work out the conceptual structure of a physical theory---or at least an appropriate one, if it is not unique---to isolate and characterize its mathematical and formal structures, and show how natural physical interpretation accrues to them from an analysis of their empirical application" (272).
-% I think that Curiel might make a similar point regarding the significance of a formulation even if it uses surplus structure or less physically significant structure.
-See section 7 argument against structural realism: Curiel argues that the Hamiltonian formulation nevertheless does not fail to represent classical systems, even if it structure is not the most natural (on his view)
-%having argued that Lagrangian mechanics has more structure ``than the theory really needs", North claims that ``we should infer that it does not exist, that it does not correspond to fundamental structure" [75].
-% Curiel on the harmonic oscillator: claims to have given an example where there is sameness of meaning (both Lagrangian and Hamiltonian formulations represent the harmonic oscillator and provide the same meaning) without sameness of structure (they're not isomorphic).
-%``we should infer that the world contains at least the amount of structure needed to formulate its fundamental dynamics in an invariant way" [80].
-% E.g., North argues that ``only certain kinds of mathematical object are appropriate for representing the structure of the world" and it is through these facts that we figure out which formulation of physics to prefer?
-[66].
-
-We agree that providing a physical interpretation of the invariants is important.
-However, some non-invariant quantities can be physically meaningful.
-For instance, the coordinates q define the units, and these are physically significant despite not being invariant.
-Units are physically significant because they allow us to assign numbers to a physical quantity.
-They are physically significant in part because they correspond to measurable quantities.
-Whereas North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North2009}, some choices of the coordinates q are less physically meaningful because they are not properly connected with measurement.
-Some arbitrary mixings of q's and p's preserve the symplectic manifold but are not as physically significant.
-For this reason, the q's are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative p's are not.
-The p's are inverses of the q's.
-Whereas symplectic geometry creates a mathematical, syntactic symmetry between the q and p variables, this is not a physically significant symmetry from the standpoint of measurement.
-Moreover: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
-(In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).
-% It follows from the way in which position quantities fix the units that $\dot{x} = v $.
-It is built into the notion of a dynamical system that half of the state variables are derivatives of the other half.
-
-In defining ``Hamiltonian mechanics" we are not privileging any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics.
-Each of these frameworks has different pragmatic costs and benefits.
-Hence, unlike North, we are comfortable acknowledging that the mathematical structure that you ``need" might not be unique.
-There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions.
-Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others.
-And in some sense, each of these three spaces involves reference to some ``unphysical" coordinates or representational means.
-Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics.
-In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.
-% One con of this extended phase space formulation: the Hamiltonian constraint is both a constraint and the generator of time evolution; hence, it must be defined in the neighborhood of the actual states, so the values it provides outside of these neighborhoods should be unphysical, yet they encode the equations of motion.
-% Should we remark on the claim that the only structure we need is whatever is sufficient to define densities locally? Hence, we need just the symplectic form but not the cotangent bundle?
-% E.g.
-in the extended phase space formulation, the equations of motion come from derivatives of the Hamiltonian constraint over the variables, but this constraint is defined on unphysical points.
-These unphysical points are plausibly a mathematical artifact, although they supply a nice tool/method for understanding physically significant aspects of the representation.
-
-\subsection{Some helpful quotations from North that could help characterize her position and relevant differences:}
-
-North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory.
-Structure comprises those quantities that remain intact when we transform one allowable coordinatization of the laws into another" [78]
-
-North definition of structure: ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" [66].
-So presumably, North would view the invariant phase space density as part of this fundamental structure.
-
-North's epistemic method for figuring out the structure: ``look at what is invariant under allowable transformations"; this tells us what the structure is [73].
-
-\section{Interpreting Lagrangian mechanics}
-
-%Discussion of gauge transformations in the context of Lagrangian vs.
-Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):
-In some sense (to be made more precise), gauge transformations are more fundamental in Lagrangian mechanics than Hamiltonian mechanics.
-The conjugate momentum $p $ functions as a dummy variable in virtue of being a gauge quantity.
-From motion equivalence, the Lagrangian has more invariant structure than the Hamiltonian [not sure what this really means].
-The Lagrangian has the advantage of being invariant under coordinate transformations, whereas the Hamiltonian is not a coordinate invariant.
-Instead, the Hamiltonian varies as the time component of a 4-co-vector.
-Nevertheless, the Lagrangian itself is physically problematic because it is a gauge-dependent quantity.
-The p variables are not of fundamental significance because we can perform a gauge transformation on the p's such that the q and $\dot{q} $ variables remain the same.
-For systems where motion equivalence holds, we are really measuring the q and $\dot{q} $ variables, indicating that this is a genuine gauge transformation of the momentum variables.
-Hence, privileging Hamiltonian mechanics does not entail, pace North, that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North2009}.
-Instead, position and kinetic momentum remain physically privileged, with conjugate momentum relegated to a gauge quantity (despite its role in the Hamiltonian formalism).
-% But there is a sense in which on the extended phase space, the Hamiltonian constraint is an invariant, yielding a coordinate-invariant way of handling systems with Hamiltonian mechanics. (?) 
-%Since the q and $\dot{q} $ variables fix the physical state, this indicates that transformed conjugate momentum is a gauge quantity. 
-%North argues that ``momentum is a truly fundamental quantity, on a par with position" [77].
-
-Claim to clarify: motion equivalence allows us to understand the gauge redundancy of the Lagrangian.
-However, it is not true that the Lagrangian has these redundancies in virtue of the system satisfying motion equivalence.
-
-Issue of the mathematical structure that is necessary/sufficient for Lagrangian mechanics:
-North footnote 20 and appendix: it is not generically true that the state space of Lagrangian mechanics comes equipped with a metric.
-Footnote 22: ``a more general, coordinate-free version" of Lagrangian mechanics ``does not make explicit use of this metric"; discussed in appendix A.
-North: ``the square of the Riemannian line element is what gives the allowable sets of coordinate transformations, the ones that preserve the Lagrangian equations of motion.
-A volume element would not suffice for this" [75].
-How does this connect with motion equivalence and its associated mathematical structure?
-
-North assumes that formulating a Lagrangian theory requires specifying a Riemannian metric; according to Curiel, this is neither necessary nor sufficient.
-Furthermore, the physical significance of a Riemannian metric in the context of a Lagrangian theory is unclear: two different Lagrangian's can have the same Lagrangian vector field and yet induce inequivalent Riemannian metrics
-
-%Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics.
-Curiel frequently appeals to naturalness and a distinction between natural vs.
-ad hoc/artificial constructions [figure out what his definition of ``naturalness" is, if he provides one].
-
-%
-
-%\bibliography{bibliography}
 \printbibliography

### 2020-11-09 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
Comments from Christine

@@
 The event $\psi$, the system was prepared in state $\psi$, will correspond to the ray (one-dimensional complex subspace) corresponding to the $\psi$ vector.
-Therefore, in general, all events in probability will corresponds to subspaces (of different dimensionality) of the Hilbder space.
+Therefore, in general, all events in probability will correspond to subspaces (of different dimensionality) of the Hilbert space.
 Projective spaces are the right tool to keep track of subspaces.
-The proof, then, consists in establishing what are the correct definitions in the space of events, map those into elements of the projective space and then, from the projective space, construct a map on the vector space directly.
+The proof, then, consists of establishing the correct definitions in the space of events, mapping those into elements of the projective space and then, from the projective space, constructing a map on the vector space directly.
 
@@
 	projective space. Each point of the projective space is called a ray,
-	because for a real vector space to would correspond to a line going through
+	because for a real vector space it would correspond to a line going through
 	the origin, with the origin removed. As we are in a complex space, the
@@
 	{Given two events $\pj{a}$ and $\pj{b}$, for example $X > x_1$ and $X < x_2$, their conjunction $\pj{a} \wedge \pj{b}$ is the event where both are true, $ x_1 < X < x_2$ in the example.
-In terms of our Hilbert spaces, both $\pj{a}$ and $\pj{b}$ corresponds to subspaces and $\pj{a} \wedge \pj{b}$ is exactly the intersection of the two, which is also a subspace.
+In terms of our Hilbert spaces, both $\pj{a}$ and $\pj{b}$ correspond to subspaces and $\pj{a} \wedge \pj{b}$ is exactly the intersection of the two, which is also a subspace.
 We should not confuse $\pj{a} \wedge \pj{b}$ with $\pj{a} | \pj{b}$: the first refers to either
@@
 	\end{defn}
-
-	{Note that the proposition $\pj{a} \wedge \pj{b}$ refers to either
-		preparation or measurement of both systems  in the respective state. 
-		The proposition $\pj{a} | \pj{b}$, instead, is the one that
-		corresponds to preparing one system in one state and measuring the
-		other system in the other state \cite{cox}. Incompatible states
-		refer to non-commuting observables.}
 
@@
 Conceptually, this ensures that the composite system contains all the properties of the components.
-Note that superselection rules or other restriction may prevent the independent preparation of all possible pairs (e.g.~two fermions cannot be jointly prepared in the same state).
-The tensor product is recovered only all pairs are compatible.
+Note that superselection rules or other restrictions may prevent the independent preparation of all possible pairs (e.g.~two fermions cannot be jointly prepared in the same state).
+The tensor product is recovered only when all pairs are compatible.
 Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components.

### 2020-11-09 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
Comments from Christine

@@
 	in classical mechanics to combine state spaces of linear systems.
-These are all maps from linear spaces to linear spaces but they differ in how the linearity of one is mapped to the linearity of the other.\footnote{For example, in the tensor product $a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct product $a \times (b+c) = a \times b + 0 \otimes c$ where $0$ is the zero vector.
+These are all maps from linear spaces to linear spaces but they differ in how the linearity of one is mapped to the linearity of the other.\footnote{For example, in the tensor product $a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct product $a \times (b+c) = a \times b + 0 \times c$ where $0$ is the zero vector.
 Also, in the tensor product $r * (a \otimes b) = (r*a) \otimes b = a \otimes (r*b)$ while in the direct product $r * (a \times b) = r*a \times r*b$.}
@@
 
-	While conceptually the derivation is straight forward, it requires the
-	use of projective space, which may be unfamiliar to the reader. This is
+	While conceptually the derivation is straightforward, it requires the
+	use of projective spaces, which may be unfamiliar to the reader. This is
 	necessary because events in probability space do not correspond to vectors in the Hilbert space, but to subspaces.
@@
 In the same way, the event $X = x_0$ corresponds to the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$.
-The map $M$ derived from the probability space is a map between states as events in probability space, and therefore will be a map from rays to rays, which is a map between the projectives spaces.
+The map $M$ derived from the probability space is a map between states as events in probability space, and therefore will be a map from rays to rays, which is a map between the projective spaces.
 The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.
@@
 	pairs) and returns a state in the projective space $\pj{\cal C}$ for
-	the composite. To visualize this geometrically, given a ray (a complex plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex plane) in $\cal C$.  Our final goal will be to find a map
+	the composite.
+To visualize the geometrical meaning of $M$ directly within the Hilbert spaces, given a ray (a complex plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex plane) in $\cal C$.
+Our final goal will be to find a map
 	$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
@@
 Now, suppose that, unknown to us, an additional second system was prepared in a pure state $\psi_{\cal B}$ and consider $p(a|\psi_{\cal A}\wedge \psi_{\cal B})$.
-In what cases are allowed to use the Born rune?
-That is, in what cases we can say $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = |\<a|\psi_{\cal A}\>|^2$?
-Only in the cases where $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = p(a|\psi_{\cal A}$ are statistically independent.
+In what cases are we allowed to use the Born rule?
+That is, in what cases can we say $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = |\<a|\psi_{\cal A}\>|^2$?
+Only in the cases where $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = p(a|\psi_{\cal A}$, where the systems  are statistically independent.
 In this sense, postulate (b) already implies  ``statistical independence'' between
@@
 	equation A|psi>=B|phi> is equivalent to $<psi|A^dag=<phi|B^dag$.
-Mathematically, there is nothing preventing one to represent state preparation and kets to represent measurements.
+Mathematically, there is nothing preventing one from representing state preparation with bras and measurements with kets.
 It is a matter of convention.

### 2020-11-16 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Edits to derivation and questions/notes for categorical equivalence issue

@@
 
-We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state (in manifold terminology, these are known as coordinates).
-Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$, yielding a real-valued vector for any given particle state, with $n$-many degrees of freedom.
-We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
+We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state.
+Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$, yielding a real-valued vector for any given particle state, with $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates,' but we will reserve that terminology for the state variables that define units.} We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
 For instance, $q$ might be position, defining units of meters.
+In general, the coordinates are the subset of the state variables that define the unit system, needed to measure properties of the system.
+
 On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function.
-This is because the distribution is a function only of the states, rather than of the particular choice of state variables, i.e.
-coordinates.
-To transform as a density means that $\rho_c$ changes as a function of the Jacobian determinant that relates the new coordinates to the old coordinates.
-To transform as a scalar means that the distribution $\rho_c$ is invariant under all coordinate transformations, so that the Jacobian determinant must in fact equal one.
-Hence, $\rho_c$ is a function only of the point $c$ and does not depend on the choice of coordinates.
-We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
+As a function defined over the state variables, the distribution transforms as a density, i.e.
+$\rho_c$ changes as a function of the Jacobian determinant that relates the new coordinates to the old coordinates.
+As a function defined over the points in phase space, the distribution transforms as a scalar, i.e.
+$\rho_c$ is invariant under all coordinate transformations.
+Hence, the Jacobian determinant must in fact equal one.
+In other words, $\rho_c$ is a function only of the point $c$ and does not depend on the choice of coordinates.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\hat{\xi}^b)$ is the distribution defined over the transformed state variables, and $\rho_c(s(\hat{\xi}^b))$ is the distribution defined over the points of phase space, where $s: \mathbb{R}^n \to \mathcal{S}$ an invertible function  mapping a particular value of the units $\hat{\xi}^b$ to a point in the particle phase space $\mathcal{S}$.} We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
 In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a density and scalar function.
-% Classical phase space (i.e.
-a symplectic manifold) is the only space that satisfies these constraints on the distribution.
+%This is because the distribution is a function only of the states, rather than of the particular choice of coordinates. 
+ % Classical phase space (i.e. a symplectic manifold) is the only space that satisfies these constraints on the distribution. 
+%(in manifold terminology, these are known as coordinates)
 
@@
 Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number.
-We will show that, in fact, there can only be a single $k^i$ coordinate in this case, conjugate to the single $q$ coordinate.
+We will show that, in fact, there can only be a single $k^i$ variable in this case, conjugate to the single $q$ coordinate.
 Assuming that the particle density $\rho_c$ is a scalar, it follows that we can arbitrarily change the unit defined by $q$ without changing the density $\rho_c$.
@@
 
-[We might be able to shorten part of the following argument simply by noting that the Jacobian determinant must equal unity, as explained verbally in a prior paragraph.
-The preceding verbal justification might be sufficient, i.e.
-that the distribution transforms as a density and a scalar.
-We could then focus on the trickier part of showing that there is only a single $k^i$ variable, based on the form of the Jacobian determinant.]
-
 Now, suppose that we change the units by transforming the coordinate $q$ to a new coordinate $\hat{q}$.
@@
 Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
+As shown above, the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one.
+Assume for contradiction that there is only one state variable, so that there are no $k^i$ variables.
+Then we would have $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right| = 1$.
+But this would mean the unit change cannot be arbitrary.
+Therefore we must have two or more state variables, so that there must be at least one $k^i$ variable.
+
+\begin{comment}
+%Longer justification for the Jacobian determinant equaling one.
+
  As defined, $\rho_c$ is a function from the particle state space $\mathcal{S}$ to the reals.
@@
 Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one.
-Assume for contradiction that there is only one state variable, so that there are no $k^i$ variables.
-Then we would have $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right| = 1$.
-But this would mean the unit change cannot be arbitrary.
-Therefore we must have two or more state variables, so that there must be at least one $k^i$ variable.
+
+\end{comment}
 
@@
 This means the volume form is given by $\omega^n$.
-Importantly, the volume form cannot be degenerate because otherwise there would be a volume in state space that could not contain states.
+Importantly, the volume form cannot be degenerate because otherwise there would be a `volume' in state space that could not contain states.
 But such a region of state space would---by definition---not be a volume.
-Therefore $\omega$ itself cannot be degenerate.
-As the degrees of freedom are independent, the number of states on a surface does not change if we translate it across independent degrees of freedom.
+Therefore the form $\omega$ itself cannot be degenerate (for otherwise, we would have states that are not counted by it).
+Next, we show that $\omega$ must be closed, using the fact that degrees of freedom are independent.
+If we translate a surface along an independent degree of freedom, this does not change the area of the surface, meaning that the number of states on that surface is invariant.
+Consider a parallelepiped, whose opposite sides have equal areas.
+Taking the integral over this closed surface (over the parallelepiped), opposite sides provide equal and opposite contributions.
+Hence, the integral over this closed volume must equal zero.
+This shows that the 2-form $\omega$ must be closed.
+Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold.
+By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
+% % the `motivation' section of the wiki page on symplectic manifolds provides a different style of derivation for the 2-form being non-degenerate and closed.
+
+%since the degrees of freedom are independent, the number of states on a surface does not change if we translate that state across independent degrees of freedom.
 If we imagine a parallelepiped, the integral over the surface must be zero (the integrals over opposite sides are equal and opposite) [explain/justify this more clearly.
 Is this coming from the two-form being antisymmetric?].
-Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold.
-By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
+%
+%if degenerate form, would have states that are not counted by the form.
 
@@
 a state variable that defines a unit.
-In fact, the trajectories can be fully described by these units and only these units [Say more to motivate this claim.
-Epistemic claim?
-Claim about measurement?].
-It follows that $x^i = q^i$, each $q^i$ is be paired with a conjugate $p_i$, and each state $\{q^i, p_i\}$ corresponds to one and only one trajectory in physical space.
+In fact, the trajectories can be fully described by these units and only these units.
+For, to specify a trajectory, we need epistemic access to a unit system, where the number of independent units is determined by the dimensionality of physical space.
+It follows that $x^i = q^i$, each $q^i$ is paired with a conjugate $p_i$, and each state $\{q^i, p_i\}$ corresponds to one and only one trajectory in physical space.
 At each point $x^i$, then, there must be infinitely many trajectories, one for each possible combination of $\{p_i\}$.
@@
 If the equations of motions were such that $v^i=v^i(q^i)$, then motion equivalence would fail as the full trajectory would be determined only by the $q^i$.
-So we must have $v^i=v^i(q^j, p_k)$.
-This relationship must be invertible or motion equivalence would fail [justify why: need to determine motion in physical space from phase space evolution].
+Hence, we must have $v^i=v^i(q^j, p_k)$, so that knowledge of the state in phase space suffices for knowledge of the trajectory.
+If motion equivalence holds, then this relationship must be invertible, meaning that knowledge of the trajectory suffices for knowledge of the phase space state.
 For systems that satisfy the motion equivalence condition, then, the following relationships hold at any given time:
@@
 In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^i$.
-Therefore, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible [Shouldn't the denominator of the Jacobian really be $\xi_j$, since it is with respect to both the q's and the p's?.
-We don't seem to justify that the Jacobian equals the Hessian until after calculating the determinant below].
-Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j}$ equals the Jacobian.
+Therefore, for fixed q, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
+Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
 Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere, which we take to be positive.
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).
+Hence, weak equivalence suffices for the existence of a Lagrangian.
 
-If we look at the whole distribution, we have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
+Next, we show that this Lagrangian describes massive particles under scalar potentials.
+If we look at the whole distribution (so that $q$ is no longer held fixed), we have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
 \begin{equation}
@@
 \end{equation}
-where $m$ is the unit conversion constant between velocity and conjugate momentum while $g_{ij}$ represents the linear dependency [of what on what?].
+where $m$ is the unit conversion constant between velocity and conjugate momentum, while $g_{ij}$ represents the linear dependency of the velocity on the momentum.
 
@@
 
+Barrett shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics are categorically equivalent provided that we define categories of models over the tangent and cotangent bundle formulations, respectively [1181-82].
+However, he also shows that if we consider the more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are not categorically equivalent [1182-83].
+This parallels the construction of Hamiltonian mechanics in Section~\ref{derivation}.
+
+\begin{enumerate}[1.)]
+
+\item Barrett footnote 16 conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
+There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" [1179].
+Does GC framework provide a definite answer on this question of inequivalence?
+
+\item Possible connection with motion equivalence principle: Sameness of empirical content (empirical equivalence) is cashed out in terms of agreeing ``on the allowable trajectories for particles in the system" [1180].
+For a Lagrangian system, this is characterized in terms of a base integral curve of the Lagrangian vector field $(X_L)^a$ that ``encodes one way that the positions of particles in a system with Lagrangian L might evolve over time", while the base integral curves of the Hamiltonian vector field perform an analogous role [1180]
+
+\begin{itemize}
+
+\item   Marsden and Abraham 1978, Theorem 3.6.2 proves that (in the hyper-regular domain) the Lagrangian vector field and the corresponding Hamiltonian vector field (obtained via Legendre transformation) have the same base integral curves.
+This means that they ``agree on how the positions of particles will evolve over time" [1181].
+Is this the same as the motion equivalence assumption?
+
+\end{itemize}
+
+\item Barrett footnote 18 wonders whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics.
+Does GC derivation show that there cannot be such a correspondingly general formulation of Lagrangian mechanics?
+
+\item Barrett also shows that the vector field formulations of Lagrangian and Hamiltonian mechanics are not categorically equivalent [1185].
+He interprets this as substantiating Curiel's claim that the ``Legendre transform does not respect the kinematical constraints of Lagrangian mechanics", namely because it shows that ``there are Lagrangians that give rise to the same Lagrangian vector field, but do not translate---via the Legendre transformation---into Hamiltonians that give rise to the same Hamiltonian vector field" [1186].
+What does this mean in the context of motion equivalence?
+
+\begin{itemize}
+
+\item Barrett interprets this situation as simply showing that the Hamiltonian vector field contains more information about the Hamiltonian than the Lagrangian vector field does about the Lagrangian.
+If two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to a arbitrary constant.
+Yet two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy (even up to a constant).
+
+\end{itemize}
+
+\item Teh and Tsementzis discuss Tulczyjew's reformulation of Hamiltonian mechanics, which shows that ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" [PDF 3].
+Does this accomplish a similar moral as the assumptions of physics framework?
+
+\item Furthermore, within this reformulation, Lagrangian submanifolds play a very important role. Does this lead to a possible privileging of Lagrangian mechanics?
+
+\end{enumerate}
+
 Possible objection to consider stemming from categorical equivalence: for systems in the hyper regular domain, does the categorical equivalence of Hamiltonian and Lagrangian mechanics show that we require the same number of assumptions to treat systems in this domain?
+
+The answer is presumably ``no", given Barrett's proof that the symplectic formulation of Hamiltonian mechanics is not categorically equivalent to the tangent bundle formulation of Lagrangian mechanics.
+
 Some relevant considerations for responding: note that the photon does not belong to the hyper regular domain.
@@
 In contrast, whenever the Lagrangian is not hyper regular, the Euler--Lagrange equations do not possess a unique solution.
-Furthermore, density plays a special role in the Hamiltonian framework, whereas it is in some sense an accident that the density and configuration space is invariant (simply because the Jacobian equals unity in Cartesian coordinates).
+Furthermore, density plays a special role in the Hamiltonian framework, whereas it is in some sense an accident that the density in configuration space is invariant (simply because the Jacobian equals unity in Cartesian coordinates).
 

### 2020-11-20 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
Changes with Lorenzo

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\documentclass[aps,prl,amsmath,amssymb,twocolumn,nofootinbib]{revtex4}
+\documentclass[aps,prl,amsmath,amssymb,twocolumn]{revtex4}
 %\documentclass[twocolumn,aps,showpacs,prl,groupedaddress]{revtex4}
@@
 \def\cc{_{\cal C}}
-%\def\comment#1{}
-\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
+\def\comment#1{}
+%\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
 \def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
@@
 
+	In this paper we derive the tensor product postulate (which, hence,
+	loses its status of postulate) from two other postulates of quantum
+	mechanics: the state postulate and the measurement postulate.
 	The tensor product postulate does not appear in all axiomatizations of
@@
 	insufficient to single out the tensor product, as other linear
-products of linear spaces exist, such as the direct product, the exterior/wedge roduct or the direct sum of vector spaces, which is used
+	products of linear spaces exist, such as the direct product, the exterior/wedge product or the direct sum of vector spaces, which is used
 	in classical mechanics to combine state spaces of linear systems.
 These are all maps from linear spaces to linear spaces but they differ in how the linearity of one is mapped to the linearity of the other.\footnote{For example, in the tensor product $a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct product $a \times (b+c) = a \times b + 0 \times c$ where $0$ is the zero vector.
-Also, in the tensor product $r * (a \otimes b) = (r*a) \otimes b = a \otimes (r*b)$ while in the direct product $r * (a \times b) = r*a \times r*b$.}
-
+Also, in the tensor product $r  (a \otimes b) = (r a) \otimes b = a \otimes (r b)$ while in the direct product $r  (a \times b) = r a \times r b$, where $r$ is a scalar.}
 	This belief may have arisen from the seminal book of Dirac
@@
 	bilinear), which is equivalent to postulating tensor products (or
-	linear functions of them). This is not an innocuous request. For
+	linear functions of them). It is not an innocuous request. For
 	example it does not hold where the composite vector space of two
@@
 	where tensor products were obtained by specifying additional physical
-	or mathematical requirements. In quantum field theory one tends to
+	or mathematical requirements. \comment{ {\em Il resto della frase da spostare in supp material} In quantum field theory one tends to
 		avoid problems connected with tensor products of infinite dimensional
@@
 		least for the infinite-dimensional case. We will consider the
-	non-relativistic setting here.
+		non-relativistic setting here.}
 
-	In this paper we derive the tensor product postulate (which, hence,
-	loses its status of postulate) from two other postulates of quantum
-	mechanics: the state postulate and the measurement postulate. We start
+	We start
 	from the natural definition of a composite system as the set of two
 	(or more) quantum systems. The composite system is therefore
-	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else},
-	it's state being fully described by all observables and only the observables
-	that fully describe its constituents. The first key insight is that first two postulates already assume that, at least in some cases, the probability distribution of the
+	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else}\comment{da mettere altrove:, namely, 
+		its state is fully described by the probability amplitudes of the system's observables and nothing else}.
+The first key insight is that the first two postulates of quantum theory (introduced below) already assume that, at least in some cases, the probability distribution of the
 	measurement outcomes of one system is independent of the other systems' (statistical independence).
-While we will see specifically how the Born implies this, it should be clear conceptually that we would not be able to talk about a system in the first place if we would not be able to characterize it independently.
+We cannot even talk about a system in the first place if we cannot characterize it independently.
 The second key insight is that, since all measurement
@@
 These insights are enough to characterize mathematically the state space of the composite: the linearity given by the Hilbert space, together with the fact that the composite system is fully described by the observables of $A$ and $B$, allows to extend the construction to the composite states that do not satisfy statistical independence (the entangled states).
-So the work consists of two interrelated efforts: a physical argument that start from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.
+So the work consists of two interrelated efforts: a physical argument that starts from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.
 
-	We will mainly focus on
+	\comment{mettere nel supplementary:	We will mainly focus on
 		kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no superselection rules or
@@
 		independent of the other systems (preparation independence).  This is
-	the only case in which the tensor product can be properly employed
-	\cite{susskind,zanardi,zanardilloyd}: the Hilbert space of composite
+		the only case in which the tensor product can be properly employed: the Hilbert space of composite
 		systems that have restrictions is {\em not} the tensor product of the
@@
 		since the tensor product formalism is very convenient and is often
-	used also in these cases, but it is something we need to bring out.
+		used also in these cases, and superselection rules are typically avoidable \cite{susskind,zanardi,zanardilloyd}. }
 
-	While conceptually the derivation is straightforward, it requires the
-	use of projective spaces, which may be unfamiliar to the reader. This is
-	necessary because events in probability space do not correspond to vectors in the Hilbert space, but to subspaces.
+	TODO: pensarci con calma: riscrivere il paragrafo seguente, magari partendo da un esempio e introducendo il concetto seguente:
+	Frase in cui diciamo che abbiamo bisogno di projective spaces perche' i vettori $|\psi\>$ non rappresentano proprieta' (direbbe Lorenzo) o events (o proposizioni) in probability space (direbbe Gabriele) ma sono i sottospazi che lo fanno.
+
+	Since quantum states are rays in Hilbert space, we need projective spaces
+	Events in probability space do not correspond to vectors in the Hilbert space, but to subspaces, which means we will need elements of projective geometry.
 If $\psi$ is a vector, the state corresponds to the ray $\pj{\psi}$, which is the subspace spanned by $\psi$.
@@
 	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
-	argument. Refer to the Supplementary Information for a more rigorous
-	proof of the \emph{same} argument.
+	argument. The supplementary material contains the mathematical details.
 
@@
 	acting on that space; (b)~The probability that a measurement of a
-	property $X$, described by the operator with spectral decomposition
+	property $X$, described by the operator with spectral decomposition \comment{Normalizzare qui e in altri posti!}
 	$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
@@
 	experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
-	ill-defined quantum systems. For example, in situations where
+	ill-defined quantum systems. \footnote{For example, in situations where
 		indistinguishability plays a role (e.g.~in QFT), one cannot consider
@@
 		the field. The electron is an excitation of the field, {\em not} a
-	(sub)system. We call this condition ``preparation independence''.  [We
+		(sub)system.} We call this condition ``preparation independence''.  [We
 	emphasize that the kinematic independence is inequivalent to dynamical
@@
 	the tensor product. We focus on pure states here: the argument can be
-	extended to mixed states using standard tools \cite{nielsenchuang}.
+	extended to mixed states using standard tools \cite{ballentine}.
 
@@
 	different states of the composite. Moreover, preparation independence
-	implies that $M$, and hence $m$, must be total maps: each subsystem of
+	implies that $M$, and hence $m$, must be total maps (condition H1): each subsystem of
 	the composite system can be independently prepared and gives rise to a
-	state of the composite (condition H1).  H1 is not sufficient
+	state of the composite.  H1 is not sufficient
 	to identify the tensor product: by itself it does not even guarantee
@@
 	the joint probability of events $a$ and $b$ given $z$ is
-	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. This essentially the definition of conditional probability: $p(b|z\wedge a) = \frac{p(a\wedge b|z)}{p(a|z)}$. Two events $a$ and $b$ are defined to be 
-	independent given $z$ if and only if $p(a\wedge b|z)=p(a|z)\:p(b|z)$.
-However, it is more instructive to define independence with $p(b|z\wedge a)=p(b|z)$.
-This makes it more evident that the probability of $b$ does not depend on conditioning on $a$.
-If we imagine $a$ and $b$ to be outcomes for particular measurements, we are saying that the probability of observing $b$ is not a function of the value measured by $a$.
-We can recover the first definition from the second.
-In fact, $p(b|z\wedge a)=\frac{p(a\wedge b|z)}{p(a|z)} =p(b|z)$ which leads to $p(a\wedge b|z)=p(a|z)\:p(b|z)$.
+	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Two events $a$ and $b$ are defined to be 
+	independent given $z$ if $p(a\wedge b|z)=p(a|z)\:p(b|z)$, which is true if and only if $p(b|z\wedge a)=p(b|z)$.
+If $a$ and $b$ are measurement outcomes, the probability of $b$ is not a function of $a$.
 
+	Riscrivere questo: scritto cosi', sembra che la regola di Born non si possa applicare sempre....
 	We note that the Born
@@
 	the inner product.
-
-	Since the square of the inner product is preserved, orthogonality and
+	This implies orthogonality and
 	the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
@@
 	In this setting, the antilinear case \eqref{h2b} corresponds to a change of convention (much like a change of sign in the symplectic form for classical mechanics) and can be ignored.
-Given a Hilbert space, in fact, we can imagine replacing all wavefunctions and all the operators with their complex-conjugate.
-These changes would effectively cancel out leaving the physics being represented unchanged  (for example, in his first papers
+Given a Hilbert space, in fact, we can imagine replacing all vectors and all the operators with their Hermitian conjugate, mapping vectors into duals $|\psi\>^\dag=\<\psi|$.
+These changes would effectively cancel out leaving the physics unchanged.
+The two equations $A|w\>=B|z\>$ and $\<w|A^\dag=\<z|B^\dag$.
+(For example, in his first papers
 	Schr\"odinger used both signs in his equation: effectively writing
 	{\em two} equivalent equations with complex-conjugate solutions
-	\cite{sch}). Another way to see this is noting that, given any equation 
-	in terms of operators acting on ket vectors, one can write an
-	equivalent equation in terms of bra vectors by simply taking the
-	Hermitian conjugate of both sides of the first equation. Namely, the
-	equation A|psi>=B|phi> is equivalent to $<psi|A^dag=<phi|B^dag$.
-Mathematically, there is nothing preventing one from representing state preparation with bras and measurements with kets.
-It is a matter of convention.
-We can now repeat the same analysis for the second
+	\cite{sch}). We can repeat the same analysis for the second
 	argument of $m$ to conclude that it is a bilinear map, condition (H2).
@@
 	by some superselection rules) can consider the same system as composed
-	of multiple subsystems. A typical example \cite{tellerbook} comes from
+	of multiple subsystems. \comment{Supplementary material? A typical example \cite{tellerbook} comes from
 		quantum field theory. It is customary in basically all quantum optics
@@
 		there is a ``primitive thisness'', which is captured in the definition
-	of system.
+		of system.}
 

### 2020-11-20 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
Adding proof from referee sketch

@@
 		\item \ref{prop_totality} + \ref{prop_bilinearity} + \ref{prop_basis} $\Rightarrow$ \ref{theo}: the composition map is the tensor product.
+
 	\end{enumerate}
+
+	\subsection{Addendum}
+
+	The above proof relies only on the independent preparation of subsystem and not their measurement.
+However, during the review process, an anonymous referee contributed a sketch for a proof that shows very directly that the state and measurement postulates also imply independence of measurements.
+The insight is that the mixture created by all possible measurement outcomes on $B$ must behave overall as a pure state on $A$.
+Since pure states are extreme points, this can only happen if every measurement outcome on $B$ leaves $A$ in a pure state, which makes the probability factorize.
+
+	\begin{prop}
+		Let $\pj{\psi}$ and $\pj{\phi}$ be two preparations for $A$ and $B$.
+Let $\pj{a}$ and $\pj{b}$ two measurements on the respective systems.
+Then $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi}) P(\pj{b} | \pj{\phi})$.
+	\end{prop}
+	\begin{proof}
+		Consider $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi})$.
+We can imagine first performing the measurement on $B$ and then conditioning the result on $A$.
+We have $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{b}, \pj{\psi} \wedge \pj{\phi}) P(\pj{b} | \pj{\psi} \wedge \pj{\phi})$ where by ``$\pj{b}, \pj{\psi} \wedge \pj{\phi}$'' we mean that the systems were prepared in $\pj{\psi}$ and $\pj{\phi}$ respectively and $\pj{b}$ was measured.
+Any preparation on $A$ can be expressed with a mixed state, and any measurement on $B$ depends only on $\phi$, so we have $P(\pj{a} | \pj{b}, \pj{\psi} \wedge \pj{\phi}) P(\pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \rho(\pj{b}, \pj{\psi}, \pj{\phi})) P(\pj{b} | \pj{\phi})$.
+We also must have $\sum_b P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi})$.
+
+		Putting it all together, we have $P(\pj{a} | \pj{\psi}) = \sum_b P(\pj{a} | \rho(\pj{b}, \pj{\psi}, \pj{\phi})) P(\pj{b} | \pj{\phi})$.
+Which means $|\psi\>\<\psi| = \sum_b P(\pj{b} | \pj{\phi}) \rho(\pj{b}, \pj{\psi}, \pj{\phi})$.
+But the only way that a mixture of mixed states can be equal to a pure state is if all mixed states are the same pure state.
+Therefore $\rho(\pj{b}, \pj{\psi}, \pj{\phi}) = |\psi\>\<\psi|$ for all $b$ and $\phi$.
+We finally have $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \rho(\pj{b}, \pj{\psi}, \pj{\phi})) P(\pj{b} | \pj{\phi}) = P(\pj{a} | \pj{\psi}) P(\pj{b} | \pj{\phi})$.
+	\end{proof}
 

### 2020-11-23 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Added derivation of stationary action principle from Hamiltonian mechanics (section 6.1)

@@
 \usepackage{amsmath, hyperref, enumerate, comment}
-\usepackage{amsfonts}
+\usepackage{amsfonts, esint}
 \usepackage{graphicx}
+\usepackage{mathrsfs} %for script font
 
@@
 
+\newcommand{\diff}{\mathop{}\!\mathrm{d}} %defines exterior derivative 
+\renewcommand{\div}[1]{\gv{\nabla} \cdot #1} % for divergence
+\newcommand{\curl}[1]{\gv{\nabla} \times #1} % for curl
+\renewcommand{\vector}[1]{\ensuremath{\vec{#1}}} % redefines vector to vec
+\newcommand{\integral}{\int}
+
 \begin{document}
-
 %% testing from desktop
@@
 
+\subsection{A geometric interpretation of the action principle}
+
+Hamilton's principle of stationary action serves as a foundational principle for Lagrangian mechanics.
+According to this principle, infinitesimal variations of the action $ \mathscr{A}$ along the system's trajectory $\gamma $ equal zero: $\delta  \mathscr{A}[\gamma] = 0 $.
+Although central to Lagrangian mechanics, we can actually derive the stationary action principle using Hamiltonian mechanics.
+Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system's motion always advances in time (so that $dt/dt = 1$).
+
+Knowing that the system is deterministic and reversible, it follows that possible trajectories of the system are never pushed toward each other or pulled apart [try to make this more precise].
+In other words, the flux of trajectories through any closed surface is always zero.
+Formally, this simply means that the displacement of the states in time is divergenceless: $\nabla \cdot \vector{S} = 0 $.
+Here, $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ defines a vector field in the extended phase space $(q, p , t ) $.
+$\vector{S}$ characterizes the displacement of states in time.
+In particular, it is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and parameter of the trajectory.
+
+Since $\vector{S}$ is divergenceless, there exists another vector field $-\vector{\theta}$ such that $\vector{S} = curl(-\vector{\theta} )$.\footnote{For reasons that will become clear, we choose $-\vector{\theta}$ rather than $\vector{\theta}$.} As a vector field in the extended phase space, $\vector{\theta}$ can be written in terms of its components $\vector{\theta} = (\theta_q, \theta_p, \theta_t) $.
+Since the curl of the gradient of a scalar function is always zero, the displacement vector $\vector{S}$ is invariant under the gauge transformation $\vector{\theta'} =\vector{\theta} + grad(f)$, where $f$ is an arbitrary scalar function.
+Hence, we can choose $f$ such that $\partial f/ \partial p = -\theta_p $.
+Without loss of generality, we can thereby redefine $\vector{\theta}$ as $(\theta_q, 0, \theta_t) $.
+
+Assuming that the motion of the system always advances in time, the time component of the displacement, $S_t = dt/ dt = 1$.
+This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $.
+In components, this means that $S_t = - (\frac{\partial}{\partial q} \cdot \theta_p - \frac{\partial}{\partial p} \cdot \theta_q) = - (\frac{\partial}{\partial q} \cdot 0 - \frac{\partial}{\partial p} \cdot \theta_q) = \frac{\partial q}{\partial p}$.
+Hence, $S_t = 1$ entails that $\frac{\partial q}{\partial p} = 1$.
+Integrating this expression, we find that $\theta_q = p$ up to addition by an arbitrary constant function $g(q)$.
+Again using the gauge freedom of $\vector{S}$ we can set $g(q)=0$, so that $\theta_q = p$.
+Thus, so far, we have shown that $\vector{\theta} = (p, 0,\theta_t) $.
+
+Since the system is deterministic and reversible, it satisfies Hamilton's equations: $\frac{\partial H}{\partial p} =\frac{dq}{dt} \equiv S_q$ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt} \equiv S_p $.
+Additionally, since $\vector{S} = -curl(\vector{\theta}) $, we know that $S_q = -\frac{\partial \theta_t}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}$.
+Integrating the first equation for $\theta_t$ gives $-H$ up to an arbitrary constant function of q, while the second gives $\theta_t = -H$ up to an arbitrary constant function of p.
+For consistency, these two constant functions must vanish, so that $\theta_t = -H$ [check argument here; do we actually need to appeal to the gauge freedom of S again?].
+Hence, it follows that we can take $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
+
+At this point, we return to the action $ \mathscr{A}$, defined as the integral of the Lagrangian over a trajectory beginning at time $t_0$ and ending at time $t_1$: $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt$.
+Provided that the Legendre transform is well-defined (i.e.
+the motion equivalence assumption is satisfied), then $H = p \dot{q} - L$.
+Hence, we can reexpress the action as $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$.
+Multiplying the Hamiltonian by $dt/dt = 1$ and adding a vanishing $0 \cdot \frac{dp}{dt}$, we obtain $ \mathscr{A} = \integral^{t_1}_{t_0} [p \frac{dq}{dt} + 0 \cdot \frac{dp}{dt} - H(q, p, t) \frac{dt}{dt} ]dt$.
+We notice that the integrand simply equals the inner product of $\vector{\theta}$ with $\vector{S} $, entailing that the Lagrangian $L =\vector{\theta} \cdot \vector{S} $.
+Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$.
+Changing variables, we take the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.
+To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
+
+By definition, a variation over the integral $\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ consists in taking the difference between the integrand and a corresponding integral over a nearby trajectory $\gamma'$: $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} - \integral_{\gamma'} \vector{\theta} \cdot d\vector{\gamma}$ [is the measure of the latter integral given according to $\gamma'$?].
+Since the trajectories $\gamma$ and $\gamma'$ have the same endpoints, they define a closed-curve $C$, enclosing a region $\Sigma $.
+Hence, by Stoke's theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \integral_{C} [\vector{\theta} + \vector{\theta}] \cdot d\vector{\gamma} = \oiint \limits_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$.
+Recalling now that 1) the $- curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$ and 2),  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $\vector{\theta}$ is tangent to [part of?] the surface $\Sigma $.
+Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
+This establishes Hamilton's principle of stationary action, which we have derived from Hamiltonian mechanics (deterministic and reversible evolution), using the Legendre transform, with the further weak assumption that the motion always advances in time.
+
 %Discussion of gauge transformations in the context of Lagrangian vs.
 Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):
+
+\subsection{Gauge redundancy in Lagrangian mechanics}
 

### 2020-11-25 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
Another pass to the tensor article

@@
 Requirement 2 ensures that it does not contain properties that are orthogonal to all the components' properties, i.e.~that the composite system contains {\em only} the components.
+Violation of the second requirement would mean that some composite states would not define all the properties of the subsystems.
+That is, while the systems $A$ and $B$ by itself would define an observable $O_A$ and $O_B$, when put grouped together all values would be assigned zero probability; those observables no longer exist.
+In this case, the nature of the system would have changed so radically we would no longer call it a composite system.
 	\end{defn}

### 2020-11-25 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
Another pass to the tensor article

@@
 	(or more) quantum systems. The composite system is therefore
-	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else}\comment{da mettere altrove:, namely, 
-		its state is fully described by the probability amplitudes of the system's observables and nothing else}.
-The first key insight is that the first two postulates of quantum theory (introduced below) already assume that, at least in some cases, the probability distribution of the
-	measurement outcomes of one system is independent of the other systems' (statistical independence).
-We cannot even talk about a system in the first place if we cannot characterize it independently.
+	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else}.\footnote{Note that this does not exclude entangled states since they too are characterized by the observables and only the observables of the constituents.} The first key insight is that the first two postulates of quantum theory (introduced below) already assume that the preparation of one system is independent from the preparation of another (statistical independence).
+In fact, we cannot even talk about a system in the first place if we cannot characterize it independently.
 The second key insight is that, since all measurement
@@
 
-	TODO: pensarci con calma: riscrivere il paragrafo seguente, magari partendo da un esempio e introducendo il concetto seguente:
-	Frase in cui diciamo che abbiamo bisogno di projective spaces perche' i vettori $|\psi\>$ non rappresentano proprieta' (direbbe Lorenzo) o events (o proposizioni) in probability space (direbbe Gabriele) ma sono i sottospazi che lo fanno.
-
-	Since quantum states are rays in Hilbert space, we need projective spaces
-	Events in probability space do not correspond to vectors in the Hilbert space, but to subspaces, which means we will need elements of projective geometry.
-If $\psi$ is a vector, the state corresponds to the ray $\pj{\psi}$, which is the subspace spanned by $\psi$.
-Note that ``ray'' can be misleading since $\pj{\psi}$ is really a complex { \em plane }.
-In the same way, the event $X = x_0$ corresponds to the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$.
-The map $M$ derived from the probability space is a map between states as events in probability space, and therefore will be a map from rays to rays, which is a map between the projective spaces.
+	To make these arguments precise, we will need to use elements of projective geometry.
+	If $\psi$ is a vector in our Hilbert space, the pure states is represented by the ray $\pj{\psi}$, which is the subspace spanned by $\psi$.
+This includes not only vectors with arbitrary phase, but also arbitrary modulus.
+The term ``ray'' can therefore be misleading since $\pj{\psi}$ is really a complex { \em plane }.
+In the same way, constraining the observable $X$ to a particular value $x_0$ means identifying the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$.
+As we need to map states and observables between subsystems and composite system, the map $M$ derived from the probability space is really a map between subspaces, between the projective spaces.
 The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.
@@
 			property''}. \commentg{We may have to revise to be more clear.
-			Where is this used?} } As discussed, we limit ourselves to
+			Where is this used?} }
+
+	Note that we limit ourselves to
 	kinemati\-cal\-ly-inde\-pen\-dent systems, where all state vectors
 	$|\psi\>$ in the system's Hilbert space $\cal H$ describe a valid
-	state, {\em unconditioned on anything else}. In particular, we note
+	state, {\em unconditioned on anything else}.
+We call this condition ``preparation independence'' and it should be noted that the tensor product applies only in this case.
+For example, the composite system of two electrons is not the tensor product, rather the anti-symmetrized tensor product, precisely because the second electron cannot be prepared in the same state of the first.
+We note
 	that restrictions due to superselection rules arise either from
@@
 	experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
-	ill-defined quantum systems. \footnote{For example, in situations where
-		indistinguishability plays a role (e.g.~in QFT), one cannot consider
-		an electron as a quantum system: in this case the quantum system is
-		the field. The electron is an excitation of the field, {\em not} a
-		(sub)system.} We call this condition ``preparation independence''.  [We
+	ill-defined quantum systems. In the example above, the field is the proper quantum system and the electrons are its excitations. \footnote{We
 	emphasize that the kinematic independence is inequivalent to dynamical
@@
 	Here we will not consider dynamical evolution, which is contained in
-	postulate (d).]
+	postulate (d).}
 
@@
 	the joint probability of events $a$ and $b$ given $z$ is
-	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$. Two events $a$ and $b$ are defined to be 
-	independent given $z$ if $p(a\wedge b|z)=p(a|z)\:p(b|z)$, which is true if and only if $p(b|z\wedge a)=p(b|z)$.
-If $a$ and $b$ are measurement outcomes, the probability of $b$ is not a function of $a$.
+	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$.
+Consider $p(a \wedge b | \psi \wedge b)$ which represents the probability of measuring $a$ on system $A$ and $b$ on system $B$ having prepared system $A$ in $\psi$ and system $B$ in $b$.
+We have $p(a \wedge b | \psi \wedge b) = p(a | \psi \wedge b \wedge b) p(b | \psi \wedge b) = p(a | \psi \wedge b ) p(b | \psi \wedge b)$.
+The Born rule tells us that $p(a | \psi \wedge b) = |\<a|\psi\>|^2$ and that $p(b | \psi \wedge b) = |\<b|b\>|^2 = 1$, where $a$, $b$ and $\psi$ are assumed to be normalized.
+We have:
+	\begin{align}
+	p(a\wedge b|\psi \wedge b)&=p(a|\psi)\\
+	p(a\wedge b|a \wedge \phi)&=p(b|\phi)
+	\end{align}
+	In other words, since the probability for a measurement on one system depends only on its pure state, the Born rule automatically assume that the measurement of one system is independent from the preparation of the other.
+We call this property ``statistical independence''.\footnote{One can also prove that the measurements on the components are independent as well, but we only strictly need preparation.}
 
-	Riscrivere questo: scritto cosi', sembra che la regola di Born non si possa applicare sempre....
-	We note that the Born
-	probability formula in postulate (b) is of the form $p(a|\psi_{\cal A}$ and contains only quantities of the
-	system (the pure state $\psi_{\cal A}$ and the observable's eigenstates $a$).
-Now, suppose that, unknown to us, an additional second system was prepared in a pure state $\psi_{\cal B}$ and consider $p(a|\psi_{\cal A}\wedge \psi_{\cal B})$.
-In what cases are we allowed to use the Born rule?
-That is, in what cases can we say $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = |\<a|\psi_{\cal A}\>|^2$?
-Only in the cases where $p(a|\psi_{\cal A}\wedge \psi_{\cal B}) = p(a|\psi_{\cal A}$, where the systems  are statistically independent.
-In this sense, postulate (b) already implies  ``statistical independence'' between
-	different systems if they are prepared independently. Namely, the
-	outcome probability of one is independent of any properties of any
-	other in the absence of any dynamical coupling and supposing
-	independent preparation. [Of course, the dynamics may couple the state
-	of the system to other systems so that the outcomes may depend on what
-	happens to other systems.  Analogously, one may prepare the system in
-	a way that depends on other systems. But we consider independent
-	preparations and do not consider dynamics, as it is sufficient for our
-	aims.] We can then formalize ``statistical independence'' of
-	independently prepared systems by saying that their Born rule
-	satisfies:
-	\begin{eqnarray}
-	p(a\wedge b|\psi_{\cal A}\wedge \psi_{\cal B})&=&p(a|\psi_{\cal A}\wedge \psi_{\cal
-		B})\:p(b|\psi_{\cal A}\wedge \psi_{\cal B})\nonumber\\&=& p(a|\psi_{\cal
-		A})\:p(b|\psi_{\cal B})
-	\labell{bindip}\;,
-	\end{eqnarray}
-	where $\psi_{\cal A}$, $\psi_{\cal B}$ represent the independent
-	preparations of the states of the systems $\cal A$, $\cal B$ and $a$,
-	$b$ the measurement outcomes of two observables on $\cal A$ and $\cal
-	B$ respectively. The first equality in equation \eqref{bindip} embodies the
-	definition of statistical independence, the second follows from the
-	form of the Born rule for independently prepared systems (the
-	probability depends only on the state and observables of each system
-	on its own).
-
-	A particular case is when one system is
-	prepared in a fixed eigenstate $|b\>$ of the observable measured on it,
-	$p(a\wedge b|\psi\wedge b)=p(a|\psi)\:p(b|b)=p(a|\psi)$, where the last equality follows from the Born rule since $\<b|b\>=1$. 
-	If we define $M_b(\pj{a}) = M(\pj{a},\pj{b})$, substituting the values of the
+	We can use this fact to characterize our map $M$.
+Note that $M(\pj{a}, \pj{b})$ will correspond to the composite state where $A$ and $B$ are prepared in the respective states.
+We can define define $M_b(\pj{a}) = M(\pj{a},\pj{b})$ and, substituting the values of the
 	probabilities from the Born rule, we have:
@@
 These changes would effectively cancel out leaving the physics unchanged.
-The two equations $A|w\>=B|z\>$ and $\<w|A^\dag=\<z|B^\dag$.
+The two equations $A|w\>=B|z\>$ and $\<w|A^\dag=\<z|B^\dag$ are equivalent.
 (For example, in his first papers

### 2020-11-30 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Edits to derivation of action principle and categorical equivalence section

--- 2020-11-16 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 \usepackage{amsmath, hyperref, enumerate, comment}
-\usepackage{amsfonts}
+\usepackage{amsfonts, esint}
 \usepackage{graphicx}
+\usepackage{mathrsfs} %for script font
 
-\usepackage[backend=biber,  style=chicago-authordate, url=false, doi=false, eprint=false]{biblatex}
+\usepackage[backend=biber, style=authoryear, url=false, doi=false, eprint=false]{biblatex}
 %\usepackage[backend=biber,  style=authoryear-icomp, url=false, doi=false, eprint=false]{biblatex}
@@
 Defending this claim does not require commitments to fundamental structure or perfectly natural properties.
+(Furthermore, the Hamiltonian framework is more general because a Hamiltonian defines a unique solution (unique evolution in phase space) even when the Hessian is zero.
+In contrast, whenever the Lagrangian is not hyper-regular, the Euler--Lagrange equations do not possess a unique solution.
+Additionally, density plays a special role in the Hamiltonian framework, whereas it is in some sense an accident that the density in configuration space is invariant (simply because the Jacobian equals unity in Cartesian coordinates).
+
+%Note that the photon does not belong to the hyper-regular domain. Furthermore, if you are not in the hyper-regular domain, then there is no longer a one-to-one map between position and velocity.
 
@@
 For instance, knowledge of the trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from its trajectory alone.
-Hence, the set of classical systems that satisfy this further physical condition is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics.
+In general, different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
+In contrast, the trajectories in phase space (encoded by the Hamiltonian vector field) uniquely define the Hamiltonian up to a constant.} Hence, the set of classical systems that satisfy this further physical condition is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics.
 On these grounds, we will argue in Section~\ref{fundamentality} that Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
+[To do: verify that the scenario of two different Lagrangians giving rise to the same trajectories in physical space with different energies, arises only when motion equivalence condition fails.
+See Curiel 2014, page 42 for his take on this kind of scenario.]
+
+% One way to motivate the motion equivalence condition: in general, the same trajectory in physical space can correspond to different trajectories in phase space: knowledge of the position and velocity is not sufficient for knowledge of the momentum (e.g.
+if you do not know the mass of the system).
+Formally, different Lagrangians can have the same Lagrangian vector field but give rise to Hamiltonians with different Hamiltonian vector fields, as Barrett notes \parencites*[1186]{Barrett2}.
+In contrast, if two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to an arbitrary constant (whereas two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy, even up to a constant)).
 
@@
 These unphysical points are plausibly a mathematical artifact, although they supply a nice tool/method for understanding physically significant aspects of the representation.
+
+% One way to motivate the motion equivalence condition: in general, the same trajectory in physical space can correspond to different trajectories in phase space: knowledge of the position and velocity is not sufficient for knowledge of the momentum (e.g.
+if you do not know the mass of the system).
+Formally, different Lagrangians can have the same Lagrangian vector field but give rise to Hamiltonians with different Hamiltonian vector fields, as Barrett notes \parencites*[1186]{Barrett2}.
+In contrast, if two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to an arbitrary constant (whereas two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy, even up to a constant)).
 
@@
 
+% [to do: make the footnote regarding connection to Marsden and Abraham more precise.
+Might want to flag that their theorem is expressed in terms of the Lagrangian vector field and corresponding Hamiltonian vector field (obtained via Legendre transformation) having the same base integral curves.
+These base integral curves encode how particles evolve over time.
+Should I also elaborate on how one can interpret motion equivalence in terms of the theories having the same empirical content?
+This is how Barrett interprets the Marsden and Abraham theorem 3.6.2]
 \begin{quotation}
 \noindent
-\textit{Motion equivalence assumption}: the motion of the system (i.e. trajectories in physical space-time) suffices to recover its dynamics (i.e. evolutions in state space) and vice-versa.
+\textit{Motion equivalence assumption}: the motion of the system (i.e.
+trajectories in physical space-time) suffices to recover its dynamics (i.e.
+evolutions in state space) and vice-versa.\footnote{ \textcites[]{Abraham1978}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time (see \textcites[1180-1181]{Barrett2} for discussion).
+Motion equivalence can be understood as the converse of this theorem: assuming that Lagrangian and Hamiltonian mechanics agree on particle evolutions, we derive that the Legendre transformation is well-defined in the hyper-regular domain.}
 \end{quotation}
@@
 
+\subsection{A geometric interpretation of the action principle}
+\label{action}
+
+% Somewhere below, emphasize that you have a unique trajectory only if the action is a minimum or maximum (if it is just stationary but not a minimum or maximum, then you do not have a unique trajectory)
+
+Hamilton's principle of stationary action serves as a foundational principle for Lagrangian mechanics.
+According to this principle, infinitesimal variations of the action $ \mathscr{A}$ along the system's trajectory $\gamma $ equal zero: $\delta  \mathscr{A}[\gamma] = 0 $.
+Although central to Lagrangian mechanics, we can actually derive the stationary action principle using Hamiltonian mechanics.
+Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition.
+We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating the generalization, see \textcites[110]{Arnold1989}.
+[To do: verify that Arnold can legitimately be read as generalizing the argument below.]}
+% Note on the generalization: cannot use curl.
+Need to use a form that measures the number of states on a surface.
+The displacement field kills this form (eats the degenerate direction that does not form degrees of freedom).
+% Something that I thought was a second assumption but turns out that this follows just from double role played by time variable: 2) the system's motion always advances in time (so that $dt/dt = 1$).
+
+Knowing that the system is deterministic and reversible, it follows that the density of states is invariant.
+This means that possible trajectories of the system are never pushed toward each other or pulled apart.
+In other words, the flux of trajectories through any closed surface is always zero.
+Formally, the displacement of the states in time is divergenceless: $\nabla \cdot \vector{S} = 0 $.
+Here, $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ defines a vector field in the extended phase space $(q, p , t ) $.
+$\vector{S}$ characterizes the displacement of states in time.
+In particular, it is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and affine parameter of the trajectory.
+
+Since $\vector{S}$ is divergenceless, there exists another vector field $-\vector{\theta}$ such that $\vector{S} = curl(-\vector{\theta} )$.\footnote{For reasons that will become clear, we choose $-\vector{\theta}$ rather than $\vector{\theta}$.} As a vector field in the extended phase space, $\vector{\theta}$ can be written in terms of its components $\vector{\theta} = (\theta_q, \theta_p, \theta_t) $.
+Since the curl of the gradient of a scalar function is always zero, the displacement vector $\vector{S}$ is invariant under the gauge transformation $\vector{\theta'} =\vector{\theta} + grad(f)$, where $f$ is an arbitrary scalar function.
+Hence, we can choose $f$ such that $\partial f/ \partial p = -\theta_p $.
+Without loss of generality, we can thereby redefine $\vector{\theta}$ as $(\theta_q, 0, \theta_t) $.
+
+Treating time $t$ as both a coordinate and an affine parameter (parameterizing the trajectory), the time component of the displacement, $S_t = dt/ dt = 1$.\footnote{Considering $dt/dt$, the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
+We could instead use a parameter $\tau$, and the derivation would remain valid.} This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $.
+In components, this means that $S_t = - (\frac{\partial}{\partial q} \cdot \theta_p - \frac{\partial}{\partial p} \cdot \theta_q) = - (\frac{\partial}{\partial q} \cdot 0 - \frac{\partial}{\partial p} \cdot \theta_q) = \frac{\partial q}{\partial p}$.
+Hence, $S_t = 1$ entails that $\frac{\partial q}{\partial p} = 1$.
+Integrating this expression, we find that $\theta_q = p$ up to addition by an arbitrary constant function $g(q)$.
+Again using the gauge freedom of $\vector{S}$ we can set $g(q)=0$, so that $\theta_q = p$.
+Thus, so far, we have shown that $\vector{\theta} = (p, 0,\theta_t) $.
+
+Next, we introduce a function $H$ and let $ \theta_t \equiv -H$.
+Hence, we write $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
+By matching components of $\vector{S}$ and $-curl(\theta)$, we show that H is the Hamiltonian for the system.
+First, by definition, $S_q \equiv \frac{dq}{dt} $ and $S_p \equiv  \frac{dp}{dt} $.
+Secondly, since $\vector{S} = -curl(\vector{\theta}) $, we calculate that $S_q = -\frac{\partial \theta_t}{\partial p} = \frac{\partial H}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}=-\frac{\partial H}{\partial q}$.
+Matching terms, we see that $\frac{\partial H}{\partial p} =\frac{dq}{dt} $ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt}$.
+These simply are Hamilton's equations, which justifies interpreting $H$ as the Hamiltonian.
+
+% Since the system is deterministic and reversible, it satisfies Hamilton's equations: $\frac{\partial H}{\partial p} =\frac{dq}{dt} \equiv S_q$ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt} \equiv S_p $.
+Additionally, since $\vector{S} = -curl(\vector{\theta}) $, we know that $S_q = -\frac{\partial \theta_t}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}$.
+Integrating the first equation for $\theta_t$ gives $-H$ up to an arbitrary constant function of q, while the second gives $\theta_t = -H$ up to an arbitrary constant function of p.
+For consistency, these two constant functions must vanish, so that $\theta_t = -H$ [check argument here; do we actually need to appeal to the gauge freedom of S again?].
+Hence, it follows that we can take $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
+
+At this point, we return to the action $ \mathscr{A}$, defined as the integral of the Lagrangian over a trajectory beginning at time $t_0$ and ending at time $t_1$: $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt$.
+Provided that the Legendre transform is well-defined (i.e.
+the motion equivalence assumption is satisfied), then $H = p \dot{q} - L$.
+Hence, we can re-express the action as $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$.
+Multiplying the Hamiltonian by $dt/dt = 1$ and adding a vanishing $0 \cdot \frac{dp}{dt}$, we obtain $ \mathscr{A} = \integral^{t_1}_{t_0} [p \frac{dq}{dt} + 0 \cdot \frac{dp}{dt} - H(q, p, t) \frac{dt}{dt} ]dt$.
+We notice that the integrand simply equals the inner product of $\vector{\theta}$ with $\vector{S} $, entailing that the Lagrangian $L =\vector{\theta} \cdot \vector{S} $.
+Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$.
+Changing variables, we take the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.\footnote{Note that $\vector{S}  dt = d\vector{\gamma}$ only on the actual trajectory in phase space, but not on other possible trajectories.} To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
+% Note on the generality of the argument: in cases where motion equivalence is not satisfied, we can still write the integrand in terms of H(q, p, t), although it would not be an integral over the Lagrangian.
+Thus, we can still express the `action' as a function of p, q, t.
+We would calculate this on a path in the extended phase space, although without motion equivalence, not all of the paths on phase space would correspond to a unique trajectory in real, physical space.
+See if Arnold has a more general name for this generalized principle of stationary action.
+
+By definition, a variation over the integral $\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ consists in taking the difference between the integrand and a corresponding integral over a nearby trajectory $\gamma'$: $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} - \integral_{\gamma'} \vector{\theta} \cdot d\vector{\gamma'}$ .
+Since the trajectories $\gamma$ and $\gamma'$ have the same endpoints, they define a closed-curve $C$, enclosing a region $\Sigma $.
+Hence, by Stoke's theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot dC = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ .
+Recalling now that 1) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and 2)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $\vector{\theta}$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
+This establishes Hamilton's principle of stationary action, which we have derived by assuming that 1) the displacement vector is divergenceless and 2) using the Legendre transform (assuming motion equivalence).
+
 %Discussion of gauge transformations in the context of Lagrangian vs.
 Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):
+
+\subsection{Gauge redundancy in Lagrangian mechanics}
 
@@
 ad hoc/artificial constructions [figure out what his definition of ``naturalness" is, if he provides one].
+
+% Helpful discussion of differences between Riemannian and symplectic manifolds from Wikipedia (https://en.wikipedia.org/wiki/Symplectomorphism): ``Unlike Riemannian manifolds, symplectic manifolds are not very rigid: Darboux's theorem shows that all symplectic manifolds of the same dimension are locally isomorphic.
+In contrast, isometries in Riemannian geometry must preserve the Riemann curvature tensor, which is thus a local invariant of the Riemannian manifold.
+Moreover, every function H on a symplectic manifold defines a Hamiltonian vector field XH, which exponentiates to a one-parameter group of Hamiltonian diffeomorphisms.
+It follows that the group of symplectomorphisms is always very large, and in particular, infinite-dimensional.
+On the other hand, the group of isometries of a Riemannian manifold is always a (finite-dimensional) Lie group.
+Moreover, Riemannian manifolds with large symmetry groups are very special, and a generic Riemannian manifold has no nontrivial symmetries."
 
@@
 
-Barrett shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics are categorically equivalent provided that we define categories of models over the tangent and cotangent bundle formulations, respectively [1181-82].
-However, he also shows that if we consider the more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are not categorically equivalent [1182-83].
+Barrett shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics are categorically equivalent provided that we define categories of models over the tangent and cotangent bundle formulations, respectively \parencites*[1181-82]{Barrett2}.
+However, he also shows that if we consider the more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are not categorically equivalent \parencites*[1182-83]{Barrett2}.
 This parallels the construction of Hamiltonian mechanics in Section~\ref{derivation}.
 
+% Not sure best place to put the following, if at all:
+Nevertheless, based on the equivalence between the tangent bundle and cotangent bundle formulations of Lagrangian and Hamiltonian mechanics, one might wonder whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics \parencites[1182]{Barrett2}.
+For, if there were, then this formulation of Lagrangian mechanics might be more general than the cotangent bundle formulation of Hamiltonian mechanics.
+And thus, there would be an analogous sense in which Lagrangian mechanics is more fundamental than Hamiltonian mechanics, undercutting our argument.
+Fortunately, it is implausible that there could be a suitable generalization of Lagrangian mechanics.
+In going from the cotangent bundle formulation of Hamiltonian mechanics to the symplectic manifold formulation, we lose some globally-defined structure.
+This is analogous to moving from the space of globally-Euclidean manifolds to the space of locally-Euclidean manifolds.
+Yet, it is not clear how Lagrangian mechanics could remain well-defined while throwing away global structure from the tangent bundle.
+For the action seems to be inherently a globally-defined quantity, as an integral defined over a trajectory.
+
+[To do: not sure if following warrants discussion:]
+Teh and Tsementzis \parencites*[]{Teh} provide another argument for the equivalence of Lagrangian and Hamiltonian mechanics within the hyper-regular domain.
+They develop this using Tulczyjew's reformulation of Hamiltonian mechanics, which shows that ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh}.
+Again, there is a worry that this equivalence result might undercut our privileging of Hamiltonian mechanics.
+However, the equivalence is again just within the hyper-regular domain, requiring that motion equivalence is satisfied.
+Furthermore, Tulczyjew's formulation relies heavily on the notion of a Lagrangian submanifold, since ``a mechanical system turns out to be a Lagrangian submanifold of an appropriately defined space" \parencites[47]{Teh}.\footnote{A submanifold $N$ of a symplectic manifold $(P, \omega) $ is Lagrangian provided that i) it has half the dimension of P and ii) the pullback of the symplectic form $\omega$ vanishes along the inclusion of N in P \parencites[48]{Teh}.} Yet importantly, Lagrangian submanifolds do not distinguish between the space of velocities and the space of momenta, i.e.
+they obscure the important physical differences between these two kinds of quantities.
+Not every constructible Lagrangian submanifold defines units in a physically significant way.\footnote{One could even construct a suitable Lagrangian submanifold by mixing position and momenta coordinates.} It is the physical significance of defining a unit system that breaks the formal (syntactic) symmetry between position and momentum coordinates.
+% [To do: verify the veracity of this discussion of Lagrangian sub manifolds]
+%Furthermore, within this reformulation, Lagrangian submanifolds play a very important role. Does this lead to a possible privileging of Lagrangian mechanics?
+
+[To do: Not sure where to put the following material, if anywhere; perhaps after a discussion of why we restrict to hyper-regular domain?:]
+In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
+There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.
+Indeed, outside the hyper-regular domain, the Euler--Lagrange equations do not have a single, unique solution.
+Hence, it is difficult to say what a Lagrangian system even is outside the hyper-regular domain.
+As an example, consider the case where the Lagrangian $L$ equals zero.
+Then for this ``system," all paths have the same action, so all paths are equally possible.\footnote{This Lagrangian corresponds to a Hamiltonian that also equals zero, entailing that $dH = 0 $.
+By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.} In contrast,  a Hamiltonian defines a unique solution (unique evolution in phase space) even when outside the hyper-regular domain (i.e.
+when the Hessian is zero).
+
+\begin{comment}
+
+Notes/questions from 11-16-20 meeting
 \begin{enumerate}[1.)]
-
-\item Barrett footnote 16 conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
-There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" [1179].
-Does GC framework provide a definite answer on this question of inequivalence?
 
@@
 Does GC derivation show that there cannot be such a correspondingly general formulation of Lagrangian mechanics?
+
+\begin{itemize}
+
+\item: GC in response: defining the Lagrangian presupposes that we have trajectories of positions over time.
+Velocity is then a derived quantity of these positions.
+So the tangent bundle seems to be intimately connected with having trajectories.
+Hence, if we eliminated reference to the tangent bundle, we would seemingly not be able to define trajectories.
+
+\item Furthermore, there is a sense in which GC derivation recovers the tangent bundle and shows why it must be well-defined: provided the evolutions are differentiable (which is needed to preserve densities), it follows that velocities are well-defined, and hence so is the tangent bundle.
+Note that the trajectories alone do not entail that velocities are well-defined, since the trajectories could in principle not be differentiable.
+
+\item Note that although the symplectic manifold already has some global structure (which corresponds to the global tangent bundle), the cotangent bundle itself is in some sense more global (and hence more specialized).
+Helpful analogy: a manifold can be locally Euclidean (which is a global future) without being globally Euclidean.
+The sense in which the tangent bundle and symplectic manifold structure are global is analogous to being locally Euclidean, whereas the global cotangent bundle structure is analogous to being globally Euclidean.
+So in going from the cotangent formulation to the symplectic formulation, we are becoming less specific (more general).
+To generalize the Lagrangian, we would seemingly need to lose some structure, but the action seems to be global (integral defined over a trajectory).
+
+\item It might be worth noting these differences, as a way of motivating the claim that there probably is not a correspondingly general formulation of Lagrangian mechanics.
+
+\end{itemize}
 
@@
 
-Possible objection to consider stemming from categorical equivalence: for systems in the hyper regular domain, does the categorical equivalence of Hamiltonian and Lagrangian mechanics show that we require the same number of assumptions to treat systems in this domain?
-
-The answer is presumably ``no", given Barrett's proof that the symplectic formulation of Hamiltonian mechanics is not categorically equivalent to the tangent bundle formulation of Lagrangian mechanics.
-
-Some relevant considerations for responding: note that the photon does not belong to the hyper regular domain.
-Furthermore, if you are not in the hyper regular domain, then there is no longer a one-to-one map between position and velocity.
-In this sense, the Hamiltonian framework is more general because it applies outside the hyper regular domain: the Hamiltonian can define a unique solution (unique evolution in phase space) even when the Hessian is zero.
-In contrast, whenever the Lagrangian is not hyper regular, the Euler--Lagrange equations do not possess a unique solution.
-Furthermore, density plays a special role in the Hamiltonian framework, whereas it is in some sense an accident that the density in configuration space is invariant (simply because the Jacobian equals unity in Cartesian coordinates).
+\end{comment}
 

### 2020-12-01 Articles/2019-QuantumComposition/Submissions/PhysLetters/response.tex
Response before Christine's review

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epsfig,amssymb,amsmath,amsthm}
%\usepackage[active]{srcltx}
%\usepackage[hypertex,linkcolor=red]{hyperref}

\usepackage{color}
\usepackage{cancel}
\usepackage{tikz-cd}

\DeclareMathOperator{\spn}{span}

\newcommand{\pj}[1] {\underbar{$#1$}}
%\newcommand{\pj}[1] {\overline{#1}}

\def\>{\rangle}
\def\<{\langle}
\def\ca{_{\cal A}}
\def\cb{_{\cal B}}
\def\cc{_{\cal C}}
\def\comment#1{}
%\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
\def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
\def\labell#1{\label{#1}}
%\def\labell#1{\label{#1}{\mbox{{\tiny #1}}}}
%\def\section#1{{\par\em #1:--- }}
\def\togli#1{}
\def\sh{\mbox{sh}}
\def\iden{\openone}

\begin{document}

Dear Dr. Robert Garisto,

thank you very much for having provided us
good quality referee reports for our paper. We appreciate that this is
not an easy task.

We believe we have overcome the criticisms pointed out by the referees
and we detail below our point-by-point response. Referee A suggested
that our paper ``could be published in PRL after substantial re-writing
and polishing of the arguments", which we have carefully done
following his/her useful suggestions, even though it took a
substantial effort to rewrite most of the paper. Referee B appreciates
our work (``There is no doubt that this paper represents some fine
mathematical work"), but questions its suitability for PRL. However,
the Referee has essentially focused on the supplementary material, in
his/her words: ``The essence of the paper, to my mind, lies in the
supplementary information." We agree with the Referee that the
supplementary information is not suitable as a paper for PRL. However,
the paper itself was carefully written to emphasize the PHYSICAL
arguments whereas the supplementary material just contains the
(nontrivial) mathematical details behind the physics. We urge the
Referee to take a fresh look at the paper with this in mind. Of course
we take full responsibility for the Referee's misconception: the
Referee is clearly very qualified, and in our previous version of the
paper we had failed to make this distinction clear. (Indeed we had
wrongly claimed that the supplementary material contained the same
material formulated in a different manner.)

Thank you for your time.

In the following we include our detailed answers to the Referees. Note 
that we will not enclose the changes made to
the paper, as we have modified most of it as requested by Referee A (removing some parts to make
space).

\section*{Referee A}

\begin{quote}
The authors address a very interesting problem that has
not received much attention, and they present an important result
about the internal logical structure of quantum theory.
\end{quote}

We thank the Referee for the positive assessment of our work.
\begin{quote}
I would say that the result is correct, but I could not
completely follow the argumentation of the article. I found the text
quite confusing and I think that it does not meet the writing
standards of PRL. In my opinion the arguments could be simplified and
expressed with simpler mathematics (see below). I think that the paper
could be published in PRL after substantial re-writing and polishing
of the arguments.
\end{quote}

We appreciate the suggestion that our paper can be published in
PRL. As suggested, we have substantially rewritten our paper following
the useful and insightful suggestions by the Referee, which we
sincerely thank for the feedback. We should remark that this was not
an easy undertaking, and we had to devote substantial effort in
rewording most of the paper. Nonetheless, we believe it was worth it,
because the argument is much clearer now.

\begin{quote}
1. Regarding 'statistical independence is not an
additional requirement: it is already contained in the measurement
postulate.' I think that this is a bit more subtle. [...]
\end{quote}

This is a great argument! We had not thought of that. The Referee is
correct, one can indeed prove the statistical independence of the
outcomes as the Referee does. And indeed the Referee is correct that
we had not considered that.

However, upon re-checking our proof very carefully, we realized that
we do not need the statistical independence of the outcomes. We just
need the statistical independence of the preparations. Nonetheless, we
believe that the argument the Referee proposes is really nice and we
respectfully ask the Referee if we can add it to our paper (citing an
anonymous Referee, or the Referee may contact us separately if he/she
wants to be properly acknowledged). For now we have added this at the end of the
supplementary material, but, if the Referee disagrees, we can remove
it and replace it with a sentence of the type: ``The statistical
independence of the outcomes is not an additional assumption, as it
can be proved independently."

The Referee's comment made us realize that we didn't discuss the
statistical independence in sufficient detail, and that is actually
one of the critical step in the demonstration (i.e. making explicit an
implicit assumption). We rewrote and expanded the part of the paper
(and the supplementary material) where we discuss this, mainly in the
first column of page 3.

\begin{quote}
2. I would say that physicists are familiar with the
notion of 'colinear'.  Hence, I would either (i) provide a definition
or (ii) express the argument with the more familiar notions 'linear'
and 'anti-linear' I would be inclined for the second (if possible).
\end{quote}

We presume that the Referee intended ``physicists are UNfamiliar".  We
agree and we also agree that we have not done enough in the text to
guide the reader in that respect. As suggested, we have explained what
``colinear" means. As far as we understand, we cannot remove this
notion from our proof, unfortunately. Indeed, the use of projective
geometry is necessary. We now have a paragraph that clearly says that
and why we need projective geometry.

This is necessary to understand the properties of the maps M and m
that we need to derive the tensor products.

\begin{quote}
3. Regarding 'the dual-space description of a quantum
system through bras is equivalent to the description using kets', what
means 'equivalent' here? Can you justify it with an argument instead
of the anecdote that 'in his first papers Schroedinger used both signs
in his equation, effectively writing two equivalent equations with
complex-conjugate solutions'?
\end{quote}

We simply meant the following. Given any equation in quantum mechanics
in terms of operators acting on ket vectors, one can write an
EQUIVALENT equation in terms of bra vectors by simply taking the
Hermitian conjugate of both sides of the first equation. Namely, the
equation $A|\psi\>=B|\phi\>$ is equivalent to $\<\psi|A^\dag=\<\phi|B^\dag$. In
terms of wave functions, it means taking the complex conjugate (which
entails a change of sign in the case of the Schroedinger equation). We
apologize this was unclear (upon re-reading that part, we agree) and
we have now specified it explicitly. We rewrote that part completely,
and we added the previous equivalence explicitly.

\begin{quote}
4. Regarding 'The function m, then, is an isomorphism over
all elements of the basis $|a_i\>|b_j\>$'. What it means to be an
'isomorphism over' some vectors?
\end{quote}

We mean that the function m is a one-to-one mapping, that maps the
elements of the basis into elements of a basis of a tensor product
space. We understand our notation could lead to confusion. We have
rewrote that part, and also changed the (admittedly confusing)
notation to clarify the whole thing.

\begin{quote}
5. Regarding 'other linear products of linear spaces
exist'. Is the direct product of two vector spaces a vector space? (I
am assuming that a composite system is also a system and hence it has
an associated Hilbert space.)
\end{quote}

Yes, a direct product of vector spaces is a vector space. You can
certainly define a Hilbert space using direct products of vectors of
two Hilbert spaces (you just need to make sure the norm doesn't
diverge). Of course, that is not the correct thing to do in quantum
mechanics. We are simply pointing out that there are many ways to map
two linear spaces to another linear space, so requiring linearity on
the spaces is not sufficient to single out the tensor product. In
fact, linearity by itself is not sufficient to warrant use of a
product to create the joint space.

We have changed that sentence to make it clearer, and have also added
a footnote to further clarify the distinction

\begin{quote}
How you define multiplication by scalars? If it is defined
distributively then this direct product becomes the direct sum of the
two spaces. And similarly with the topological product. Therefore,
these three examples of product of linear spaces are the same one.
\end{quote}

Each linear space comes with its own sum (e.g. $+ : A \times A \to A$)
and its own scalar multiplication (e.g. $* : \mathbb{R} \times A \to
A$). Distributivity is $r * (a + b) = (r*a) + (r*b)$. Each space
individually will retain that property.

If we imagine a map $m : A \times B \to C$, we can ask how can we
express $r*m(a,b)$ as a function of $m(r*a)$ and $m(r*b)$. This
differs from product to product. For the direct product, we have $r *
(a \times b) = r*a \times r*b$ which is distributive. For the tensor
product, we have $r * (a \otimes b) = r*a \otimes b = a \otimes r*b$
which makes it not distributive.

We can also ask how we can express $m(a,b+c)$ in terms of $m(a,b)$ and
$m(a,c)$. For the tensor product we have $a \otimes (b+c) = a \otimes
b + a \otimes c$ which is distributive. For the direct product we have
$a \times (b+c) = a \times b + 0 \times c$ where $0$ is the zero
vector. So the direct product is not distributive.

The direct sum and direct product are the same in the finite case but
not in the infinite case. In that case, one can only take finitely
many non-zero vectors for the direct sum, while this is not required
in the direct product.

The overall point is: there are many choices. For distributivity, one
has to decide if it's distributivity over the addition or the scalar
multiplication. That choice needs to be motivated. The exterior
product is also distributive over the addition, like the tensor
product (as it is a quotient of the tensor product). So, why the
tensor product and not the exterior product?

We added the following footnote, hoping it will succinctly highlight
the issue: For example, in the tensor product $a \otimes (b+c) = a
\otimes b + a \otimes c$ while in the direct product $a \times (b+c) =
a \times b + 0 \times c$ where $0$ is the zero vector. Also, in the
tensor product $r * (a \otimes b) = (r*a) \otimes b = a \otimes (r*b)$
while in the direct product $r * (a \times b) = r*a \times r*b$.

This also addresses the next point.

\begin{quote}
6. Regarding '(Dirac) adds the seemingly innocuous
request'. I understand that requiring the product of vectors to be
distributive is the same as requiring it to be bilinear, that is $|a\>
(|b\> + |c\>) = |a\>|b\> + |a\>|c\>$. Am I missing something?
\end{quote}

No, what the Referee states is correct. Our point is that such request
is far from innocuous, as for example it excludes (without any
physical justification whatsoever) the direct product. The requirement
of distributivity (or bilinearity) introduced by Dirac is an
ADDITIONAL assumption that is certainly not warranted and it is even
NOT natural. For example, it does not hold in classical mechanics,
where the state space of composite systems is obtained through the
direct product, and (a,b+c) is DIFFERENT from (a,b)+(a,c).

\begin{quote}
Supplementary information 1. Definition of composite
system. Why do you say `for any compatible pair' instead of `for any
pair'? Can we have a pair of systems where not all pair of states are
compatible?
\end{quote}

We can have a pair of systems where not all pairs of states are
compatible. Consider two electrons. They cannot be prepared in the
same state. Therefore the state space of the composite system should
exclude those cases. We use the exterior product (i.e. the
anti-symmetrized tensor product) in those cases. As we explicitly say
elsewhere, electrons cannot be properly be considered as quantum
subsystems (at least in the case in which indistinguishability plays a
role): the quantum system is the field, and electrons are just field
excitations (quanta).

We now made this more explicit after the definition: Requirement 1
ensures that the composite system is well defined at least when the
components are prepared independently. Conceptually, this ensures that
the composite system contains all the properties of the
components. Note that superselection rules or other restrictions may
prevent the independent preparation of all possible pairs (e.g.~two
fermions cannot be jointly prepared in the same state). The tensor
product is recovered only when all pairs are compatible.

\begin{quote}
``I would say that this high degree of generality obscures
the argument. Perhaps this case of systems with non-compatible pairs
of states can be left for a discussion after the main proof."
\end{quote}
We agree with the Referee, but we suppose that the Supplementary
information section will be only read by people who want to understand
the full details of our argument. This type of readers will not be
scared by the full generality of the argument. Instead in the main
paper we were more careful to consider first the compatible states and
then the general issues. In the Supplementary information we need to
consider the situation in full generality, e.g.  systems of multiple
fermions and bosons when considered as systems as is sometimes done
improperly (a common enough case that we have to justify why, in those
cases, the tensor product is not used directly). In the proof, we
already concentrate on the preparation independent case as we also
didn't want to make the derivation too heavy. We hope that the above
clarification explains why we believe we have to make that distinction
mathematically and why it is also interesting physically. In any case,
we have clarified this point both in the paper and the supplementary
material.

\begin{quote}
2. Regarding 'the proposition a wedge b is equivalent to
the (pure) state $M (a, b)$'. The authors do not define what it means
for a logical proposition to be equivalent to a ray? Is it necessary
to use the formalism of logical propositions to derive the tensor
product? If so, some definitions should be introduced.  Similarly,
what it means to say that a state $M (a, b)$ corresponds to an event in
probability space. I have the feeling that the same can be said in
more plain words.
\end{quote}

We agree that we haven't clearly explained the formalism and we have
now done so. We do not really need the formalism of logical
propositions, but just the logical connective ``AND" that we indicated
with the wedge symbol. That's all we need, and it's necessary for
defining the JOINT probability. We could remove the wedge symbol and
replace it with the connective "and". We tried, but the formulas
become quite heavy, so we believe that this (minimal) concession to
the formalism is useful.

We have modified the paper (and the supplementary material) in
multiple places to clarify the formalism: not only the formalism of
logical propositions, but also how these are mapped in the formalism of quantum mechanics
(e.g. rays and Hilbert subspaces).

\begin{quote}
3. Regarding `Every preparation of the composite gives a
nontrivial measurement on the components.' Do you mean that, for any
state of the composite system, local projective measurements must have
at least one outcome with non-zero probability? I think that this
phrasing is easier to accept.
\end{quote}
Yes. That is exactly what we meant. We agree that the Referee's
wording is clearer and we have replaced it in the new version of our
paper.

In conclusion, we thank the Referee for the careful reading and the
thoughtful suggestions. They have been all very valuable in enhancing
the paper.

\section*{Referee B}

\begin{quote}
This paper shows how the tensor-product assumption of
quantum theory can be derived from two standard postulates of quantum
theory--specifically, the postulates about states and
measurements--together with a set of assumptions reflecting reasonable
interpretations of the expressions 'composite system' and 'independent
systems'. The essence of the paper, to my mind, lies in the
supplementary information.
\end{quote}

We respectfully disagree with this assessment: while the supplementary
information contains the nontrivial mathematical formulation of the
ideas presented in our paper, we believe that our main contribution is
in the PHYSICS, rather than in the MATHEMATICS. Namely, we show how
some PHYSICAL ASSUMPTIONS (contained in the other postulates of
quantum mechanics) must give rise to tensor products. We have
carefully written our paper on two separate levels (actually three):
at a first level, a physicist interested in the main physical ideas
will only read the paper, and can easily skip the sketch of the proof
presented there, while still understanding the main results. At a
second level, a reader that wants more details can also follow the
sketch of the proof that we put in the main paper. Finally, a
specialist of quantum mechanics (such as, clearly, the Referee) can
get into the fine details of the argument by looking at the physical
considerations in the paper and the mathematical details in the
supplementary information. We understand that this distinction was not
clear in the first version of the paper, and take we take
responsibility for this (we ourselves had mischaracterized our
supplementary information). We urge the Referee to re-read the paper
keeping in mind that the main paper is written for a more general
audience and that mostly contains the physics, and the supplementary
information only contains the details of the mathematical proofs.
\begin{quote}
This is where one finds a precise statement of each
assumption and a precise articulation of the logical connections
between statements. Indeed, the authors' main contribution is the
articulation of this logical argument. A number of physicists may
already hold the intuition that the tensor product rule is the only
composition rule one can imagine in light of the rest of the theory;
however, making this intuition precise, as the authors have done, is
another matter.\end{quote}

We thank the Referee for this compliment. We agree: most physicists do
hold the intuition that the tensor product is the natural composition
rule for quantum mechanics, but, as the Referee suggests, this had
indeed never been shown previously, at least using the conventional
formulation of quantum mechanics as we do.

\begin{quote}There is no doubt that this paper represents some fine
mathematical work.\end{quote}

Thank you!

\begin{quote}The question is whether it meets the criteria for
publication in PRL.  As the authors point out, the tensor product
postulate has been addressed by other authors. Some authors explain
why this postulate is reasonable but do not prove in detail how it
follows from other assumptions. Within the quantum logic community,
the tensor product rule has been derived rigorously from simpler
axioms. The authors' approach is different in that they are starting
with two other standard postulates of quantum theory and thereby aim
to show that there is some redundancy in the standard postulates."
\end{quote}
We agree: this is a good characterization of our work.

\begin{quote}In this respect, the authors' approach could be compared
to that of Ref. [8], which derives the measurement postulate from
other standard postulates (including the tensor product
postulate). Ref. [8] derives not only the mathematical representation
of measurement outcomes, but also the Born rule and the general form
of the rule for determining the post-measurement state of the
system. That paper was published in Nature Communications. Another
paper, Ref. [28], which also derives the Born rule from the postulate
about states and the tensor product rule, was published many years ago
in American Journal of Physics.
\end{quote}

It is true that both our result and theirs shows that there is a certain redundancy in the postulates, even though our result is completely different from theirs. So, following the Referee's suggestion we have emphasized this.

\begin{quote}However, I think there is an important difference between
the tensor product postulate and the measurement postulate. The
measurement postulate is the central focus of the notorious
measurement problem of quantum theory, which has been addressed in
many books and papers and which many would argue is the root of all
the interpretational difficulties. (Indeed, Ref. [28] explicitly
addresses the measurement problem.) The tensor product postulate does
not have this status. So, whereas a paper deriving the measurement
postulate might well be read by any physicist with even a passing
interest in the conceptual foundations of the theory, a paper deriving
the tensor product postulate will not have the same breadth of
appeal. Also, I think it is somehow less surprising that the tensor
product postulate emerges in this way than that the measurement
postulate does. The most obvious alternative to the tensor product is
the direct sum, which one immediately sees does not capture what is
meant by 'composite system'.\end{quote}

We certainly agree with the Referee that a solution of the measurement
problem would be more interesting than our result. BUT it is quite
debatable (and actually debated) whether any of the above results
resolved the measurement problem: the general consensus is that it is
still OPEN. In contrast, our result is quite solid (and neither
Referee objects to that): we DO show that the tensor product postulate
is not a necessary postulate, and believe (more on this below) that
our result is certainly suitable for publication on PRL. Referee A
shares this belief.

\begin{quote}
As I look over the four criteria the editors of PRL have
given for determining whether a paper should be published in PRL, I do
not see that any of them clearly applies to this paper. This paper
will be read with interest by physicists who have thought carefully
about the logical structure of quantum theory, but I do not think PRL
is the right place to publish it.
\end{quote}

We believe the Referee has seen these four criteria in a too
restrictive light. We please ask the Referee to take a look at the PRL
papers that are usually accepted for publication. As a typical
example, the first paper that currently appears on the PRL web site is
titled ``Stochastic Interpolation of Sparsely Sampled Time Series via
Multipoint Fractional Brownian Bridges". We have been professional
physicists for decades, and none of us has any idea of what that even
means. (We emphasize that this is the first paper on the PRL web site
today: we did not choose a particularly cryptic one!) Instead, any
physicist (or even a physics student) that has taken a quantum
mechanics course (i.e. all physicists?) will understand our title and
abstract: they have met tensor products when dealing with multiple
systems.

In addition, we also believe that the way in which we construct
Hilbert spaces from physical statements is also relevant: we cannot
claim that this is a new result, of course (although we could not find
a paper where our same formalization is presented), but this is
typically not emphasized in existing literature. In particular, we
show how the link between events in probability and observables in
Hilbert spaces goes through the projective spaces, and how the choice
of representation, the choice of gauge (phase and modulus of the state
vectors), plays into that. So, in addition to our main result, we also
claim that popularizing this particular connection is also a strength
of our paper.

\begin{quote}
I also have a couple of specific comments. 1. I think the
authors are emphasizing too strongly the smallness of the number of
postulates they are using (e.g., in the paper's title).
\end{quote}

We used the usual quantum postulates that appear in the literature,
e.g. in the referenced literature. We have no qualms in adding
postulates if the Referee suggests that we should do it. Our title is
invariant: indeed, we could also title our paper ``The five postulates
of quantum mechanics are four" or so on for any n: ``The n postulates
of quantum mechanics are n-1". Our main result is that we can remove
one of the postulates (the tensor product one): how many we start from
is not very relevant.

\begin{quote} As I was
reading the main text, it felt as if the authors were sometimes
finding new postulates in their interpretations of the words used in
definitions, and yet they were declining to call them postulates,
possibly because to do so would contradict the paper's title. I
understand that these `smaller' assumptions can be thought of as
simply making precise the meanings of commonly used expressions-- this
fact becomes clearer in the supplementary information. But strictly speaking, these are genuine assumptions.\end{quote}

We disagree: we did not add ``new postulates". 
Let us first be precise as to what a postulate is.
If one looks carefully, all postulates of quantum mechanics links something
defined within the physics world (e.g. states, observables, composite system,
time evolution) to some mathematical object (e.g. rays in a Hilbert space, Hermitian
operators, tensor product, unitary evolution). The idea is that we are
establishing a link between two different definitions (the physical
and the mathematical) that have no a priori justification. But each
element is well defined within its realm. We are not adding new postulates in the sense that we are not adding any
new link of this kind. What we are doing is showing that the definition of 
a composite system, which is ultimately a definition (NOT a postulate) within the physics
realms, together with the links already established between state space/Hilbert
space and measurment/Born rule is enough to demonstrate that the state space of the composite
system must be the tensor product of the subsystems.

More specifically, for example, there is a fundamental difference between a statement
like ``the state of a quantum system is represented by a vector in a
Hilbert space" and something like ``the state of a composite system
will fully determine the probability distribution of the observable of
all its parts and only of its parts." The first connects two
completely distinct elements, the physical preparation of a quantum
system and the mathematical construct of a Hilbert space. There is no
a priori reason for that and we are not giving one in this paper. The
second is defining an object based on the other. That is, we do not
define composite systems in some other way and than later we
independently find that they have the property of characterizing the
subsystems.  We define the composite system to be that. There is no
assumption being made: it is just a definition, and it is the natural
definition that comes from our physical intuition.

\begin{quote} For
example, Definition I.4, part 2, which says essentially that no new
orthogonal states arise when one combines two systems, is a
substantial postulate. After all, it is not the case that no new
states at all arise when one combines two systems-- entangled states
arise, and they are new. One can imagine a world in which not only
entangled states but also new orthogonal states arise. We do not live
in that world, but one needs a postulate to rule out that world. For
the authors, that postulate is Definition I.4, part 2.
\end{quote}

That "world" would simply correspond to a different definition of a composite system. Whereas we used the natural one: the (set) union of the subsystems. 
In the standard definition, the
composite system must still define all the properties of the subsystems. A new orthogonal state would not be part of the set of eigenstates of the
operators corresponding to the observables of the subsystem.

We understand that our paper was definitely unclear on this and we have
clarified this point.

Of course, entangled states arise and that is a
trivial consequence of linearity: we start from FACTORIZED states,
those are all we need to obtain tensor products, but once tensor
products are established, then ENTANGLED states arise immediately from
linearity in the composite system's Hilbert space.

\begin{quote}
(Note that the authors' informal definition of a composite
system, which is that it consists of nothing but the two systems,
could alternatively be read as excluding entangled states. Indeed, a
reader not familiar with quantum theory might well read it that way. I
mention this just to point out the ambiguity associated with an
informal definition and to highlight the importance of the actual
mathematical postulate.)
\end{quote}

We do agree that the mathematical definitions are important to precisely
address the issue: that is why we have included all the mathematical
details in the supplementary material. However, the mathematical definitions
are not as arbitrary as the referee implies.

As the Referee correctly points out, our definition certainly does not
exclude entangled states. However, we see the Referee's point and we
agree that a reader unfamiliar with quantum theory might get confused,
so we added an explicit mention that our construction can be extended
to include entangled states, which are the ones that do not satisfy
statistical independence. We added this at the end of the paragraph,
rather than at the end of that sentence, because we did not want to
ruin the flow of the narration by introducing the concept of "states"
in that sentence (which refers to systems).

\begin{quote}
2. On page 2, postulate (a) begins, 'The state of a system
is described by a ray...'. As expressed here, this postulate actually
contradicts the authors' ultimate conclusion, which allows entangled
states of the combined system. As the authors certainly know very
well, if the combined system is in an entangled state, the state of
each component system is not described by a ray in that system's
Hilbert space. I'm not sure why the authors didn't say, 'A pure state
of a system is described by a ray...'. (The word 'pure' is missing
also in the statement of this postulate in the supplementary
information.)
\end{quote}

Definitely: the word 'pure' is missing! We apologize for this
oversight and thank the Referee for pointing it out. We certainly
should have added the specification that the postulate (as expressed
in our paper) refers to pure states. Also Referee A implicitly pointed
this out. We have now corrected it.

In conclusion, we thank the Referee B for his/her insightful
observations and helpful comments. We believe that his/her assessment
on the suitability of our paper for PRL stems for an overly
restrictive interpretation of the publishing criteria, as discussed
above, and from giving more weight to the supplementary information
than to the paper. The paper mainly contains the physical ideas (which
we believe are our main contribution) whereas the supplementary
information contains only the mathematical details, as the Referee
also pointed out.

In this respect, before it was possible to attach supplementary
material to PRLs, we used to write a PRL to give the main idea and
then an adjoining PRA or PRD paper to provide the technical
details. If the Referee suggests so, we can do it also in this case
(by transforming the supplementary material into a PRA
paper). However, we would rather not do that.

--------------------

In closing, we were very pleasurably surprised by the quality of both
reports and we thank both Referees for their time they spent in
evaluating our paper.

Sincerely,

The Authors

\end{document}

### 2020-12-01 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
After Lorenzo's review

@@
 	\title{The four postulates of quantum mechanics are three}
-	\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^{2}$ and Christine A. Aidala$^1$
-	}\affiliation{\vbox{1.~Physics Department, University of Michigan, 450 Church Street,
-			Ann Arbor, MI 48109-1040,
+	\author{Gabriele Carcassi$^1$, Lorenzo Maccone$^{2}$ and Christine A.
+		Aidala$^1$ }\affiliation{\vbox{1.~Physics Department, University of
+			Michigan, 450 Church Street, Ann Arbor, MI 48109-1040,
 			United States}\\
-		\vbox{2.~Dip.~Fisica and INFN Sez.\ Pavia, University
-			of Pavia, via Bassi 6, I-27100 Pavia, Italy}}
+		\vbox{2.~Dip.~Fisica and INFN Sez.\ Pavia, University of Pavia, via
+			Bassi 6, I-27100 Pavia, Italy}}
 	\begin{abstract}
@@
 	loses its status of postulate) from two other postulates of quantum
-	mechanics: the state postulate and the measurement postulate.
-	The tensor product postulate does not appear in all axiomatizations of
+	mechanics: the state postulate and the measurement postulate.  The
+	tensor product postulate does not appear in all axiomatizations of
 	quantum mechanics: it has even been called ``postulate 0'' in some
@@
 	insufficient to single out the tensor product, as other linear
-	products of linear spaces exist, such as the direct product, the exterior/wedge product or the direct sum of vector spaces, which is used
-	in classical mechanics to combine state spaces of linear systems.
-These are all maps from linear spaces to linear spaces but they differ in how the linearity of one is mapped to the linearity of the other.\footnote{For example, in the tensor product $a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct product $a \times (b+c) = a \times b + 0 \times c$ where $0$ is the zero vector.
-Also, in the tensor product $r  (a \otimes b) = (r a) \otimes b = a \otimes (r b)$ while in the direct product $r  (a \times b) = r a \times r b$, where $r$ is a scalar.}
-	This belief may have arisen from the seminal book of Dirac
-	\cite{diracbook}, who introduces tensor products (Chap.~20) by
+	products of linear spaces exist, such as the direct product, the
+	exterior/wedge product or the direct sum of vector spaces, which is
+	used in classical mechanics to combine state spaces of linear systems.
+	These are all maps from linear spaces to linear spaces but they differ
+	in how the linearity of one is mapped to the linearity of the others
+	\footnote{For example, in the tensor product
+		$a \otimes (b+c) = a \otimes b + a \otimes c$ while in the direct
+		product $a \times (b+c) = a \times b + 0 \times c$ where $0$ is the
+		zero vector. Also, in the tensor product
+		$r (a \otimes b) = (r a) \otimes b = a \otimes (r b)$ while in the
+		direct product $r (a \times b) = r a \times r b$, where $r$ is a
+		scalar.}.  This belief may have arisen from the seminal book of
+	Dirac \cite{diracbook}, who introduces tensor products (Chap.~20) by
 	appealing to linearity. However, he adds the seemingly innocuous
@@
 	bilinear), which is equivalent to postulating tensor products (or
-	linear functions of them). It is not an innocuous request. For
-	example it does not hold where the composite vector space of two
-	linear spaces is described by the direct product, e.g.~in classical
-	mechanics, for two strings of a guitar: it is not distributive.
-	[General classical systems, not only linear ones, are also composed
-	through the direct product.] Of course, Dirac is not constructing an
-	axiomatic formulation, so his `sleight of hand' can be forgiven. In
-	contrast, von Neumann (\cite{vonneumannbook} Chap.~VI.2, also
-	\cite{jauch}) introduces tensor products by noticing that this is a
-	natural choice in the position representation of wave mechanics (where
-	they were introduced in \cite{weyl,epr}), and then {\em explicitly
-		postulates} them in general: ``This rule of transformation is
-	correct in any case for the coordinate and momentum operators [...]
-	and it conforms with the [observable axiom and its linearity
-	principles], we therefore postulate them generally.''
-	\cite{vonneumannbook}.  More mathematical or conceptually-oriented
-	modern formulations (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang})
-	introduce this postulate explicitly.  An interesting alternative is
-	provided in \cite{ballentinebook,ballentinepaper}: after introducing
-	tensor products, Ballentine verifies a posteriori that they give the
-	correct laws of composition of probabilities. Similarly, Peres uses
+	linear functions of them). It is not an innocuous request. For example
+	it does not hold where the composite vector space of two linear spaces
+	is described by the direct product, e.g.~in classical mechanics, for
+	two strings of a guitar: it is not distributive.  [General classical
+	systems, not only linear ones, are also composed through the direct
+	product.] Of course, Dirac is not constructing an axiomatic
+	formulation, so his `sleight of hand' can be forgiven. In contrast,
+	von Neumann (\cite{vonneumannbook} Chap.~VI.2, also \cite{jauch})
+	introduces tensor products by noticing that this is a natural choice
+	in the position representation of wave mechanics (where they were
+	introduced in \cite{weyl,epr}), and then {\em explicitly postulates}
+	them in general: ``This rule of transformation is correct in any case
+	for the coordinate and momentum operators [...]  and it conforms with
+	the [observable axiom and its linearity principles], we therefore
+	postulate them generally.''  \cite{vonneumannbook}.  More mathematical
+	or conceptually-oriented modern formulations
+	(e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}) introduce this
+	postulate explicitly.  An interesting alternative is provided in
+	\cite{ballentinebook,ballentinepaper}: after introducing tensor
+	products, Ballentine verifies a posteriori that they give the correct
+	laws of composition of probabilities. Similarly, Peres uses
 	relativistic locality \cite{peres}. While these procedures seemingly
@@
 	which (in contrast to what is done here) are not connected to the
-	other postulates. Similar approaches appear in \cite{marmo,aerts}
-	where tensor products were obtained by specifying additional physical
-	or mathematical requirements. \comment{ {\em Il resto della frase da spostare in supp material} In quantum field theory one tends to
-		avoid problems connected with tensor products of infinite dimensional
-		spaces by focusing on algebraic commutation structures,
-		e.g.~\cite{giddins,roos}.  In particular, the recent MIP*=RE result
-		\cite{mipre} implies that, in infinite dimensions, the tensor product
-		is strictly less computationally powerful than the commutation
-		structures, emphasizing the difference among these two structures, at
-		least for the infinite-dimensional case. We will consider the
-		non-relativistic setting here.}
+	other postulates.  In \cite{marmo,aerts} tensor products were obtained
+	by specifying additional physical or mathematical requirements.
+	\comment{ {\em Il resto della frase da spostare in supp material} In
+		quantum field theory one tends to avoid problems connected with
+		tensor products of infinite dimensional spaces by focusing on
+		algebraic commutation structures, e.g.~\cite{giddins,roos}.  In
+		particular, the recent MIP*=RE result \cite{mipre} implies that, in
+		infinite dimensions, the tensor product is strictly less
+		computationally powerful than the commutation structures,
+		emphasizing the difference among these two structures, at least for
+		the infinite-dimensional case. We will consider the non-relativistic
+		setting here.}
 
-	We start
-	from the natural definition of a composite system as the set of two
-	(or more) quantum systems. The composite system is therefore
-	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing else}.\footnote{Note that this does not exclude entangled states since they too are characterized by the observables and only the observables of the constituents.} The first key insight is that the first two postulates of quantum theory (introduced below) already assume that the preparation of one system is independent from the preparation of another (statistical independence).
-In fact, we cannot even talk about a system in the first place if we cannot characterize it independently.
-The second key insight is that, since all measurement
-	outcomes and all states of the subsystems plus the composite live in the same probability space, we will automatically have a map $M$ that for a pair of subsystems' states gives us the composite state that corresponds to the statistically independent case.
-These insights are enough to characterize mathematically the state space of the composite: the linearity given by the Hilbert space, together with the fact that the composite system is fully described by the observables of $A$ and $B$, allows to extend the construction to the composite states that do not satisfy statistical independence (the entangled states).
-So the work consists of two interrelated efforts: a physical argument that starts from the first two postulates and leads to the necessary existence of the composition map $M$ and its properties together with a formal argument that shows how $M$ leads to the tensor product.
+	We start from the natural definition of a composite system as the set
+	of two (or more) quantum systems. The composite system is therefore
+	made of system $A$ {\em and} (joined with) system $B$ and {\em nothing
+		else}. The first key insight is that the first two postulates of
+	quantum theory (introduced below) already assume that the preparation
+	of one system is independent from the preparation of another
+	(statistical independence). In fact, we cannot even talk about a
+	system in the first place if we cannot characterize it independently.
+	The second key insight is that, since all measurement outcomes and all
+	states of the subsystems plus the composite live in the same
+	probability space, we will automatically have a map $M$ that for a
+	pair of subsystems' states gives us the composite state that
+	corresponds to the statistically independent case. These insights are
+	enough to characterize mathematically the state space of the
+	composite: the linearity given by the Hilbert space, together with the
+	fact that the composite system is fully described by the observables
+	of $A$ and $B$, allows to extend the construction to the composite
+	states that do not satisfy statistical independence (the entangled
+	states). So the work consists of two interrelated efforts: a physical
+	argument that starts from the first two postulates and leads to the
+	necessary existence of the composition map $M$ and its properties
+	together with a formal argument that shows how $M$ leads to the tensor
+	product.
 
 	\comment{mettere nel supplementary: We will mainly focus on
-		kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no superselection rules or
-		other restrictions to the state space are present: it is possible to
-		prepare each subsystem of a composite system in a state that is
-		independent of the other systems (preparation independence).  This is
-		the only case in which the tensor product can be properly employed: the Hilbert space of composite
-		systems that have restrictions is {\em not} the tensor product of the
-		component spaces, but a subspace of it (e.g.~the anti-symmetric
-		subspace for fermions). Typically this is ignored in the literature,
-		since the tensor product formalism is very convenient and is often
-		used also in these cases, and superselection rules are typically avoidable \cite{susskind,zanardi,zanardilloyd}. }
+		kine\-mati\-cal\-ly-inde\-pen\-dent systems, namely no
+		superselection rules or other restrictions to the state space are
+		present: it is possible to prepare each subsystem of a composite
+		system in a state that is independent of the other systems
+		(preparation independence).  This is the only case in which the
+		tensor product can be properly employed: the Hilbert space of
+		composite systems that have restrictions is {\em not} the tensor
+		product of the component spaces, but a subspace of it (e.g.~the
+		anti-symmetric subspace for fermions). Typically this is ignored in
+		the literature, since the tensor product formalism is very
+		convenient and is often used also in these cases, and superselection
+		rules are typically avoidable \cite{susskind,zanardi,zanardilloyd}.
+	}
 
-	To make these arguments precise, we will need to use elements of projective geometry.
-	If $\psi$ is a vector in our Hilbert space, the pure states is represented by the ray $\pj{\psi}$, which is the subspace spanned by $\psi$.
-This includes not only vectors with arbitrary phase, but also arbitrary modulus.
-The term ``ray'' can therefore be misleading since $\pj{\psi}$ is really a complex { \em plane }.
-In the same way, constraining the observable $X$ to a particular value $x_0$ means identifying the subspace spanned by all eigenstates such that $X | \psi \> = x_0 |\psi\>$.
-As we need to map states and observables between subsystems and composite system, the map $M$ derived from the probability space is really a map between subspaces, between the projective spaces.
-The technical work, then, is to show that the map $M$ on the projective spaces corresponds to a map $m$ on the vector spaces and that $m$ can be taken to be, without loss of generality, the tensor product.
+	This map $M$ acts on the state spaces of the subsystems. Each pure
+	state is identified by a ray $\pj{\psi}$, a subspace of the system's
+	Hilbert space spanned by all vectors $\psi$ differing by their
+	(nonzero) modulus and phase: a two-dimensional subspace (a plane). In
+	the same way, constraining the observable $X$ to a particular outcome
+	value $x_0$ means identifying the subspace spanned by all
+	non-normalized eigenvectors $|x_0\>$ of arbitrary phase such that
+	$X | x_0 \> = x_0 |x_0\>$. As we need to map states and observables
+	between subsystems and composite system, the map $M$ derived from the
+	probability space is really a map between subspaces, namely the
+	projective spaces. We then use projective geometry to show that the
+	map $M$ on the projective spaces corresponds to a map $m$ on the
+	vector spaces which is the tensor product.
 
-	This is done by showing that preparation independence and
-	statistical independence imply three conditions on the map $m$:
+	This is done by showing that preparation independence and statistical
+	independence of the systems imply three conditions on the map $m$:
 	(H1)~totality: the map is defined on all states of the subsystems;
@@
 	Fig.~\ref{f:fig}. The rest of the paper contains the sketch of this
-	argument. The supplementary material contains the mathematical details.
+	argument, including all the physical arguments outlined above. The
+	supplementary material contains the mathematical details.
 
@@
 		%\includegraphics[width=\linewidth, trim={0.2in 5.8in 2.5in 0.1in}, clip=true]{fig.eps}
-		\caption{Schematic depiction of the logical implications used
-			in this paper. FTPG stands for ``Fundamental Theorem of Projective Geometry''.  \label{f:fig}}\end{figure}
+		\caption{Schematic depiction of the logical implications used in
+			this paper. FTPG stands for ``Fundamental Theorem of Projective
+			Geometry''.  \label{f:fig}}\end{figure}
 
-	We start with the axiomatization of quantum mechanics based on the following
-	postulates (e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state of a
-	system is described by a ray $\pj{\psi}$ corresponding to a set of
-	non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
+	We start with the axiomatization of quantum mechanics based on the
+	following postulates
+	(e.g.~\cite{ozawa,masanes,wootters,nielsenchuang}): (a)~The pure state
+	of a system is described by a ray $\pj{\psi}$ corresponding to a set
+	of non-zero vectors $|\psi\>$ in a complex Hilbert space, and the
 	system's observable properties are described by self-adjoint operators
 	acting on that space; (b)~The probability that a measurement of a
-	property $X$, described by the operator with spectral decomposition \comment{Normalizzare qui e in altri posti!}
-	$\sum_{x,i}x|x_i\>\<x_i|$ ($i$ a degeneracy index), returns a value
-	$x$ given that the system is in state $\pj{\psi}$ is
-	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2$ (Born rule). (c)~The state
-	space of a composite system is given by the tensor product of the
-	spaces of the component systems; (d)~The time evolution of an isolated
-	system is described by a unitary operator acting on a vector
-	representing the system state, $|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or,
-	equivalently, by the Schr\"odinger equation. The rest of quantum
-	theory can be derived from these axioms. While some axiomatizations
-	introduce further postulates, we will be using only (a) and (b) to
-	derive (c), so the above are sufficient for our aims.
+	property $X$, described by the operator with spectral decomposition
+	$\sum_{x,i}x|x_i\>\<x_i|/|\<x_i|x_i\>|$ ($i$ a degeneracy index),
+	returns a value $x$ given that the system is in state $\pj{\psi}$ is
+	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2/|\<x_i|x_i\>\<\psi|\psi\>|$ (Born
+	rule). (c)~The state space of a composite system is given by the
+	tensor product of the spaces of the component systems; (d)~The time
+	evolution of an isolated system is described by a unitary operator
+	acting on a vector representing the system state,
+	$|\psi({t})\>=U_{t}|\psi({t}=0)\>$ or, equivalently, by the
+	Schr\"odinger equation. The rest of quantum theory can be derived from
+	these axioms. While some axiomatizations introduce further postulates,
+	we will be using only (a) and (b) to derive (c), so the above are
+	sufficient to our aims.
 
@@
 
-	Note that we limit ourselves to
-	kinemati\-cal\-ly-inde\-pen\-dent systems, where all state vectors
-	$|\psi\>$ in the system's Hilbert space $\cal H$ describe a valid
-	state, {\em unconditioned on anything else}.
-We call this condition ``preparation independence'' and it should be noted that the tensor product applies only in this case.
-For example, the composite system of two electrons is not the tensor product, rather the anti-symmetrized tensor product, precisely because the second electron cannot be prepared in the same state of the first.
-We note
-	that restrictions due to superselection rules arise either from
-	practical (not fundamental) limitations on the actions of the
+	Note that we limit ourselves to kinemati\-cal\-ly-inde\-pen\-dent
+	systems, where all state vectors $|\psi\>$ in the system's Hilbert
+	space $\cal H$ describe a valid state, {\em unconditioned on anything
+		else}. We call this condition ``preparation independence'' and it
+	should be noted that the tensor product applies only in this case. For
+	example, the composite system of two electrons is not the tensor
+	product, rather the anti-symmetrized tensor product, precisely because
+	the second electron cannot be prepared in the same state of the first.
+	We note that restrictions due to superselection rules arise either
+	from practical (not fundamental) limitations on the actions of the
 	experimenter \cite{susskind,zanardi,zanardilloyd} or from the use of
-	ill-defined quantum systems. In the example above, the field is the proper quantum system and the electrons are its excitations. \footnote{We
-	emphasize that the kinematic independence is inequivalent to dynamical
-	independence (or isolation).  Indeed if two systems interact, their
-	interaction may lead to dynamical restrictions in the state spaces.
-	Here we will not consider dynamical evolution, which is contained in
-	postulate (d).}
+	ill-defined quantum systems. In the example above, the field is the
+	proper quantum system and the electrons are its excitations.
+	\footnote{We emphasize that the kinematic independence is inequivalent
+		to dynamical independence (or isolation).  Indeed if two systems
+		interact, their interaction may lead to dynamical restrictions in
+		the state spaces.  Here we will not consider dynamical evolution,
+		which is contained in postulate (d).}
 
@@
 	composite system. Since states are defined by postulate (a) as rays in
-	the respective Hilbert spaces, there must exist a map $M : \pj{\cal A}
-	\times \pj{\cal B } \to \pj{\cal C}$ that takes a pair of states for
-	the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$ represent the
-	projective spaces and the Cartesian product is the set of all possible
-	pairs) and returns a state in the projective space $\pj{\cal C}$ for
-	the composite.
-To visualize the geometrical meaning of $M$ directly within the Hilbert spaces, given a ray (a complex plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex plane) in $\cal C$.
-Our final goal will be to find a map
+	the respective Hilbert spaces, there must exist a map
+	$M : \pj{\cal A} \times \pj{\cal B } \to \pj{\cal C}$ that takes a
+	pair of states for the subsystems ($\pj{\cal A}$ and $\pj{\cal B }$
+	represent the projective spaces and the Cartesian product is the set
+	of all possible pairs) and returns a state in the projective space
+	$\pj{\cal C}$ for the composite. To visualize the geometrical meaning
+	of $M$ directly within the Hilbert spaces, given a ray (a complex
+	plane) in each of $\cal A$ and $\cal B$, $M$ returns a ray (a complex
+	plane) in $\cal C$.  Our final goal will be to find a map
 	$m:{\cal A}\times{\cal B}\to{\cal C}$ that acts on vectors in the
-	Hilbert spaces $\cal A$, $\cal B$ and $\cal C$ consistently with $M$. Namely,
-	$\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign indicates the
-	elements in the projective space.
-Again geometrically, $m$ takes a vector in each of $\cal A$ and $\cal B$, and returns a vector in $\cal C$ and we want this to be consistent with $M$ such that vectors picked from the same rays will return vectors in the same ray.
-We will prove that the map $m$ is
-	the tensor product. We focus on pure states here: the argument can be
-	extended to mixed states using standard tools \cite{ballentine}.
+	Hilbert spaces $\cal A$, $\cal B$ and $\cal C$ consistently with $M$.
+	Namely, $\pj{m(a,b)}=M(\pj{a},\pj{b})$ where the underline sign
+	indicates the elements in the projective space. Again geometrically,
+	$m$ takes a vector in each of $\cal A$ and $\cal B$, and returns a
+	vector in $\cal C$ and we want this to be consistent with $M$ such
+	that vectors picked from the same rays will return vectors in the same
+	ray. We will prove that the map $m$ is the tensor product. We focus on
+	pure states here: the argument can be extended to mixed states using
+	standard tools \cite{ballentinebook}.
 
@@
 	different states of the composite. Moreover, preparation independence
-	implies that $M$, and hence $m$, must be total maps (condition H1): each subsystem of
-	the composite system can be independently prepared and gives rise to a
-	state of the composite.  H1 is not sufficient
-	to identify the tensor product: by itself it does not even guarantee
-	that the map $m$ is linear.
+	implies that $M$, and hence $m$, must be total maps (condition H1):
+	each subsystem of the composite system can be independently prepared
+	and gives rise to a state of the composite.  H1 is not sufficient to
+	identify the tensor product: by itself it does not even guarantee that
+	the map $m$ is linear.
 
-	Postulate (b) contains the connection between quantum mechanics
-	and probability theory. It must then implicitly contain the
-	axiomatization of probability, e.g.~see
-	\cite{ballentinepaper,ballentinebook,cox}. One of the axioms of
-	probability theory (axiom 4 in \cite{ballentinepaper}) asserts that
-	the joint probability of events $a$ and $b$ given $z$ is
-	$p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$.
-Consider $p(a \wedge b | \psi \wedge b)$ which represents the probability of measuring $a$ on system $A$ and $b$ on system $B$ having prepared system $A$ in $\psi$ and system $B$ in $b$.
-We have $p(a \wedge b | \psi \wedge b) = p(a | \psi \wedge b \wedge b) p(b | \psi \wedge b) = p(a | \psi \wedge b ) p(b | \psi \wedge b)$.
-The Born rule tells us that $p(a | \psi \wedge b) = |\<a|\psi\>|^2$ and that $p(b | \psi \wedge b) = |\<b|b\>|^2 = 1$, where $a$, $b$ and $\psi$ are assumed to be normalized.
-We have:
+	Postulate (b) contains the connection between quantum mechanics and
+	probability theory. It must then implicitly contain the axiomatization
+	of probability, e.g.~see \cite{ballentinepaper,ballentinebook,cox}.
+	One of the axioms of probability theory (axiom 4 in
+	\cite{ballentinepaper}) asserts that the joint probability events $a$
+	and $b$ given $z$ is $p(a\wedge b|z)=p(a|z)\:p(b|z\wedge a)$.
+	Consider $p(a \wedge b | \psi \wedge b)$ which represents the
+	probability of measurement outcomes $a$ on system $A$ and $b$ on
+	system $B$ given that system $A$ was prepared in $\psi$ and system $B$
+	in $b$.  We have
+	$p(a \wedge b | \psi \wedge b) = p(a | \psi \wedge b \wedge b) p(b |
+	\psi \wedge b) = p(a | \psi \wedge b ) p(b | \psi \wedge b)$. The Born
+	rule tells us that $p(a | \psi \wedge b) = |\<a|\psi\>|^2$ and that
+	$p(b | \psi \wedge b) = |\<b|b\>|^2 = 1$, where $|a\>$, $|b\>$ are the
+	normalized eigenstates relative to outcomes $a$ and $b$, and $|\psi\>$
+	is the normalized state vector. We have:
 	\begin{align}
@@
 	\end{align}
-	In other words, since the probability for a measurement on one system depends only on its pure state, the Born rule automatically assume that the measurement of one system is independent from the preparation of the other.
-We call this property ``statistical independence''.\footnote{One can also prove that the measurements on the components are independent as well, but we only strictly need preparation.}
-
-	We can use this fact to characterize our map $M$.
-Note that $M(\pj{a}, \pj{b})$ will correspond to the composite state where $A$ and $B$ are prepared in the respective states.
-We can define define $M_b(\pj{a}) = M(\pj{a},\pj{b})$ and, substituting the values of the
-	probabilities from the Born rule, we have:
+	In other words, since the probability for a measurement on one system
+	depends only on its pure state, the Born rule requires that the
+	measurement of one system is independent from the preparation of the
+	other. We call this property ``statistical
+	independence'' \footnote{One can also prove that the measurements on
+		the components are independent as well (see Supplementary material),
+		but we only strictly need preparation here.}.  It characterizes the
+	map $M$, since $M(\pj{a}, \pj{b})$ corresponds to the composite state
+	where $A$ and $B$ are prepared in the states $|a\>$ and $|b\>$.
+	Define $M_b(\pj{a}) = M(\pj{a},\pj{b})$. From the Born rule we find
 	\begin{align} &\Big|\Big\<M\left(\pj{a},\pj{b}\right)\Big|M\left(\pj{\psi},\pj{b}\right)\Big\>_{\cal C}\Big|^2
@@
 	\end{align}
-	where the first and second terms contain the inner product in the composite
-	space $\cal C$. [This is not a new assumption: it follows from the
-	measurement postulate (b) for the composite system.] This means that,
-	when one subsystem is prepared in an eigenstate of what is measured
-	there, the state space of the other is mapped preserving the square of
-	the inner product.
-	This implies orthogonality and
-	the hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
-	colinear transformation by definition.
-Geometrically, recall that $M_b$ maps rays to rays.
-The fact that $M_b$ is colinear means that it also maps higher order subspaces to higher order subspaces (lines to lines, planes to planes, and so on) while preserving inclusion (if a line is within a plane, the mapped line will be within the mapped plane).
-In this case, the fundamental
-	theorem of projective geometry \cite{fun} applies, which tells us that
-	a unique semi-linear map $m_b$ that acts on the vectors exists in accordance with $M_b$.
-	Moreover, conservation of probability further constrains it to be
-	either linear or antilinear. This tells us that the corresponding $m$
-	is either linear or antilinear in the first argument. Namely, if equation
+	where the first and second terms contain the inner product in the
+	composite space $\cal C$. [This is not a new assumption: it follows
+	from the measurement postulate (b) for the composite system.] This
+	means that, when one subsystem is prepared in an eigenstate of what is
+	measured there, the state space of the other is mapped preserving the
+	square of the inner product.  This implies orthogonality and the
+	hierarchy of subspaces are preserved through $M_b$, making $M_b$ a
+	colinear transformation by definition. Geometrically, recall that
+	$M_b$ maps rays to rays. The fact that $M_b$ is colinear means that it
+	also maps higher order subspaces to higher order subspaces (lines to
+	lines, planes to planes, and so on) while preserving inclusion (if a
+	line is within a plane, the mapped line will be within the mapped
+	plane). In this case, the fundamental theorem of projective geometry
+	\cite{fun} applies, which tells us that a unique semi-linear map $m_b$
+	that acts on the vectors exists in accordance with $M_b$.  Moreover,
+	conservation of probability further constrains it to be either linear
+	or antilinear. This tells us that the corresponding $m$ is either
+	linear or antilinear in the first argument. Namely, if equation
 	\eqref{questa} holds, then
@@
 	\end{align}
-	In this setting, the antilinear case \eqref{h2b} corresponds to a change of convention (much like a change of sign in the symplectic form for classical mechanics) and can be ignored.
-Given a Hilbert space, in fact, we can imagine replacing all vectors and all the operators with their Hermitian conjugate, mapping vectors into duals $|\psi\>^\dag=\<\psi|$.
-These changes would effectively cancel out leaving the physics unchanged.
-The two equations $A|w\>=B|z\>$ and $\<w|A^\dag=\<z|B^\dag$ are equivalent.
-(For example, in his first papers
-	Schr\"odinger used both signs in his equation: effectively writing
-	{\em two} equivalent equations with complex-conjugate solutions
-	\cite{sch}). We can repeat the same analysis for the second
+	In this setting, the antilinear case \eqref{h2b} corresponds to a
+	change of convention (much like a change of sign in the symplectic
+	form for classical mechanics) and can be ignored.  Given a Hilbert
+	space, in fact, we can imagine replacing all vectors and all the
+	operators with their Hermitian conjugate, mapping vectors into duals
+	$|\psi\>^\dag=\<\psi|$. These changes would effectively cancel out
+	leaving the physics unchanged: the two equations $A|w\>=B|z\>$ and
+	$\<w|A^\dag=\<z|B^\dag$ are equivalent.  (For example, in his first
+	papers Schr\"odinger used both signs in his equation: effectively
+	writing {\em two} equivalent equations with complex-conjugate
+	solutions \cite{sch}). We can repeat the same analysis for the second
 	argument of $m$ to conclude that it is a bilinear map, condition (H2).
 
-	The last condition (H3) follows directly from the definition of a
-	composite system. Since it is composed {\em only} of the component
-	systems, for any state $c$ of the composite system, we must find at least one pair $(a, b)$ such that $p(a\wedge b | c)\neq 0$. It follows that the map $m$ is span-surjective: namely the
-	span of the map applied to all states in the component systems spans
-	the composite system state space. In other words, the composite does
-	not contain states that are totally independent of (i.e.~orthogonal
-	to) the states of the components.
+	The last condition, span surjectivity (H3), follows directly from the
+	definition of a composite system. Since it is composed {\em only} of
+	the component systems, for any state $c$ of the composite system, we
+	must find at least one pair $|a\>$, $|b\>$ such that
+	$p(a\wedge b | c)\neq 0$.  Span-surjectivity follows: namely the span
+	of the map applied to all states in the component systems spans the
+	composite system state space. In other words, the composite does not
+	contain states that are totally independent of (i.e.~orthogonal to)
+	the states of the components.
 
@@
 	tensor product. More precisely, given a total, span-surjective,
-	bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ that preserves the
-	square of the inner product, we find that $\cal C $ is equivalent to
+	bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ that maps the
+	Hilbert spaces $\cal A$, $\cal B$ of the components into the Hilbert
+	space $\cal C$ of the composite and that preserves the square of the
+	inner product, we find that $\cal C $ is equivalent to
 	$\cal A \otimes \cal B $ and that $m=\otimes$.
 
-	Proof. Step 1: the bases of the component systems are mapped to a basis of
-	the composite system. Because of totality property (H1) and because
-	the square of the inner product is preserved, we can conclude that,
-	given two orthonormal bases $\{|a_i\>\}\in{\cal A}$ and
+	Proof. Step 1: the bases of the component systems are mapped to a
+	basis of the composite system. Because of totality property (H1) and
+	because the square of the inner product is preserved, we can conclude
+	that, given two orthonormal bases $\{|a_i\>\}\in{\cal A}$ and
 	$\{|b_j\>\}\in{\cal B}$,
@@
 	tensor product ${\cal A}\otimes{\cal B}$ and the associated bilinear
-	map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the property
-	than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$ factors
-	through $T$ uniquely.  This means that there exists a {\em unique}
-	$\hat m$, dependent on $m$, such that $\hat m \circ T=m$.  In other
-	words, the following diagram commutes:
+	map $T : \cal A \times \cal B \to {\cal A}\otimes{\cal B}$ have the
+	property than any bilinear map $m:{\cal A}\times{\cal B}\to{\cal C}$
+	factors through $T$ uniquely.  This means that there exists a {\em
+		unique} $\hat i$, dependent on $m$, such that $\hat i \circ T=m$.
+	In other words, the following diagram commutes:
 	\begin{center}
-		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{m}"] \\
+		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{i}"] \\
 			& \mathcal{C}
@@
 	\end{center}
-	Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is
-	a bilinear operator (property H2), thanks to the universal property of
-	the tensor product we can find a unique linear operator $\hat{m} :
-	\mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that $m(a, b) =
-	\hat m(a \otimes b)$. The set $\{ \hat m(a_i\otimes b_j)$ with
-	$|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and ${\cal B}\}$
-	forms a basis for $\cal C$, since $\hat m(a_i\otimes b_j)=m(a_i,b_j)$
-	and we have shown above that the latter is a basis.  Thus,
+	Since $m : \mathcal{A} \times \mathcal{B} \to \mathcal{C}$ is a
+	bilinear operator (property H2), thanks to the universal property of
+	the tensor product we can find a unique linear operator
+	$\hat{i} : \mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that
+	$m(a, b) = \hat i(a \otimes b)$. The set $\{ \hat i(a_i\otimes b_j)$
+	with $|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and
+	${\cal B}\}$ forms a basis for $\cal C$, since
+	$\hat i(a_i\otimes b_j)=m(a_i,b_j)$ and we have shown above that the
+	latter is a basis.  Thus,
 	\begin{align} 
-	&\<\hat m(a_i\otimes b_j)|\hat m(a_k\otimes b_\ell)
+	&\<\hat i(a_i\otimes b_j)|\hat i(a_k\otimes b_\ell)
 	\>_{\mathcal{C}}=\<m(a_i,
@@
 	where we used the orthonormality of the bases and the fact that
-	$|a_i\otimes b_j\>$ is a basis of the tensor product space ${\cal A}\otimes{\cal B}$.
-Since the function $\hat{m}$ is a linear function that maps an orthonormal basis of ${\cal A}\otimes{\cal B}$ to an orthonormal basis of $\cal C$, $\hat{m}$ is a an isomorphism (a bijection that preserves the mathematical structure) between ${\cal A}\otimes{\cal B}$ and $\cal C$.
-As ${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert spaces, they are mathematically equivalent and we can directly use the tensor product to represent the composite state space.
-This means that the map $m : \cal A \times \cal B \to \cal C$ is equivalent to the map $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$ in the sense that $m \circ \hat m^{-1} = \otimes$.$\square$
+	$|a_i\otimes b_j\>$ is a basis of the tensor product space
+	${\cal A}\otimes{\cal B}$. Since the function $\hat{i}$ is a linear
+	function that maps an orthonormal basis of ${\cal A}\otimes{\cal B}$
+	to an orthonormal basis of $\cal C$, $\hat{i}$ is a an isomorphism (a
+	bijection that preserves the mathematical structure) between
+	${\cal A}\otimes{\cal B}$ and $\cal C$. As
+	${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert
+	spaces, they are mathematically equivalent: we can loosely say that
+	$\hat i$ is the identity. So we can directly use the
+	tensor product to represent the composite state space. This means that
+	the map $m : \cal A \times \cal B \to \cal C$ is equivalent to the map
+	$\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$ in the
+	sense that $m \circ \hat i^{-1} = \otimes$.$\square$
 
-	\togli{\comment{Commento finale da aggiungere?  E' interessante perche' tutto
-			cio' ci dice che non otteniamo necessariamente il prodotto tensore,
-			ma il prodotto tensore modulo una fase locale che e' fisicamente
-			irrilevante. Viene da chiedersi se c'e' una qualche notazione
-			(evidentemente piu' generale di quella di Dirac) che ci permetta di
-			eliminare questa ambiguita' fisicamente irrilevante... La notazione
-			di Dirac gia' elimina la necessita' di avere una rappresentazione
-			per descrivere gli stati. Pero' evidentemente non elimina del tutto
-			la ambiguita' di rappresentazione perche' un cambio di coordinate
-			sia sugli stati che sugli operatori mi lascia invariata la fisica,
-			ma non lascia invariata la notazione di Dirac...  Probabilmente la
-			notazione che elimina l'ambiguita' e' quella che utilizza le matrici
-			densita' invece dei vettori di stato (vedi Ozawa \cite{ozawa} e
-			Holevo): le matrici densita' normalizzate sono phase-independent:
-			$\rho=|\psi\>\<\psi|$. Attenzione: l'interpretazione di una matrice
-			mista $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$ come ``il sistema e' nello
-			stato $|\psi_i\>$ con probabilita' $p_i$'' e' una CONSEGUENZA della
-			regola di Born, quindi nella formalizzazione del postulato degli
-			stati in termini di matrici densita', questa interpretazione non
-			puo' apparire perche' e' una conseguenza di un altro postulato.}}
+	\togli{\comment{Commento finale da aggiungere?  E' interessante
+			perche' tutto cio' ci dice che non otteniamo necessariamente il
+			prodotto tensore, ma il prodotto tensore modulo una fase locale
+			che e' fisicamente irrilevante. Viene da chiedersi se c'e' una
+			qualche notazione (evidentemente piu' generale di quella di Dirac)
+			che ci permetta di eliminare questa ambiguita' fisicamente
+			irrilevante... La notazione di Dirac gia' elimina la necessita' di
+			avere una rappresentazione per descrivere gli stati. Pero'
+			evidentemente non elimina del tutto la ambiguita' di
+			rappresentazione perche' un cambio di coordinate sia sugli stati
+			che sugli operatori mi lascia invariata la fisica, ma non lascia
+			invariata la notazione di Dirac...  Probabilmente la notazione che
+			elimina l'ambiguita' e' quella che utilizza le matrici densita'
+			invece dei vettori di stato (vedi Ozawa \cite{ozawa} e Holevo): le
+			matrici densita' normalizzate sono phase-independent:
+			$\rho=|\psi\>\<\psi|$. Attenzione: l'interpretazione di una
+			matrice mista $\rho=\sum_ip_i|\psi_i\>\<\psi_i|$ come ``il sistema
+			e' nello stato $|\psi_i\>$ con probabilita' $p_i$'' e' una
+			CONSEGUENZA della regola di Born, quindi nella formalizzazione del
+			postulato degli stati in termini di matrici densita', questa
+			interpretazione non puo' apparire perche' e' una conseguenza di un
+			altro postulato.}}
 
@@
 	universal property of the tensor product, which uniquely characterizes
-	it. First we show that the bilinear map $m$ maps subsystems' bases
+	it. In step 1 we show that the bilinear map $m$ maps subsystems' bases
 	into the composite system basis. We also know that there exists a
 	tensor product map $ {T}=\otimes$ that can compose the vectors in
-	$\cal A$ and $\cal B$.  Then we use the universal property: since $m$
-	is a bilinear map, we are assured that there exists a unique $\hat m$
-	such that $\hat m \circ T=m$. Since we show that $\hat m$ is an
-	isomorphism, then $\hat{m}$ bijectively maps vectors in $\cal C$ onto
-	vectors in the tensor product space. Namely $m={T}=\otimes$.
+	$\cal A$ and $\cal B$.  In step 2 we use the universal property: since
+	$m$ is a bilinear map, we are assured that there exists a unique
+	$\hat i$ such that $\hat i \circ T=m$. Since we show that $\hat i$ is
+	an isomorphism, then $\hat{i}$ bijectively maps vectors in $\cal C$
+	onto vectors in the tensor product space. Namely $m={T}=\otimes$.
 
@@
 	by some superselection rules) can consider the same system as composed
-	of multiple subsystems. \comment{Supplementary material? A typical example \cite{tellerbook} comes from
-		quantum field theory. It is customary in basically all quantum optics
-		literature to treat different modes of the radiation field (e.g.~the
-		output of two lasers) as independent systems composed through the
-		tensor product.  Clearly the electromagnetic field is a single system
-		and an agent who is able to access an optical mode that is a linear
-		combination of the two will give a quantum description for it that
-		cannot easily accommodate tensor products. Similarly, an agent can
-		consider two electrons as two systems, joined with the tensor product,
-		whenever they are distinguishable for all practical purposes (e.g.~the
+	of multiple subsystems. \comment{Supplementary material? A typical
+		example \cite{tellerbook} comes from quantum field theory. It is
+		customary in basically all quantum optics literature to treat
+		different modes of the radiation field (e.g.~the output of two
+		lasers) as independent systems composed through the tensor product.
+		Clearly the electromagnetic field is a single system and an agent
+		who is able to access an optical mode that is a linear combination
+		of the two will give a quantum description for it that cannot easily
+		accommodate tensor products. Similarly, an agent can consider two
+		electrons as two systems, joined with the tensor product, whenever
+		they are distinguishable for all practical purposes (e.g.~the
 		electrons are in widely separated physical locations). Yet, in
@@
 		excitations (states) of the fields. In the words of Teller
-		(\cite{tellerbook}, pg.22), tensor products can be safely used only if
-		there is a ``primitive thisness'', which is captured in the definition
-		of system.}
+		(\cite{tellerbook}, pg.22), tensor products can be safely used only
+		if there is a ``primitive thisness'', which is captured in the
+		definition of system.}
 
@@
 	funding from the Attract project through the Eu Horizon 2020 research
-	and innovation programme under grant agreement No 777222.
-G.C.
-and C.A.A.
-would like to thank M.
-J.
-Greenfield for reviewing the mathematical details and acknowledge funding from the MCubed program of the University
-	of Michigan.
+	and innovation programme under grant agreement No 777222. G.C. and
+	C.A.A. would like to thank M. J. Greenfield for reviewing the
+	mathematical details and acknowledge funding from the MCubed program
+	of the University of Michigan.
 
@@
 		(2016).
-		\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum Computation
-		and Quantum Information (Cambridge University Press, Cambridge,
-		2000).
+		\bibitem{nielsenchuang}M. A. Nielsen and I. L. Chuang, Quantum
+		Computation and Quantum Information (Cambridge University Press,
+		Cambridge, 2000).
 		\bibitem{ballentinebook}L.E. Ballentine, Quantum Mechanics, a modern
@@
 		joint system, Helv. Phys. Acta {\bf 51}, 661 (1979).
-		\bibitem{giddins}S.B.~Giddings, Hilbert space structure in quantum
-		gravity: an algebraic perspective. J. High Energ. Phys. 2015, 1
-		(2015).% https://doi.org/10.1007/JHEP12(2015)099
-		\bibitem{roos}H. Roos, Independece of Local Algebras in Quantum Field
-		Theory, Comm.  Math. Phys. {\bf 16}, 238 (1970).
-		\bibitem{mipre}Z. Ji, A. Natarajan, T. Vidick, J. Wright, H. Yuen, MIP*=RE, arXiv:2001.04383 (2020). %https://quantumfrontiers.com/2020/03/01/the-shape-of-mip-re/
 		\bibitem{susskind}Y. Aharonov, L. Susskind, Charge superselection
@@
 		press, 1961).
-		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc (1957)
-		\bibitem{sch}E. Schr\"odinger, Annalen der Physik {\bf 102}, 81 (1926); English translation in E. Schr\"odinger, {\em Collected papers on Wave Mechanics} (Blackie \& Son, London, 1928).
-		\bibitem{tellerbook}P. Teller, An Interpretive Introduction to Quantum
-		Field Theory (Princeton Univ. Press, 1997).  
-		\bibitem{teller}M. Redhead, P. Teller, Particles, Particle Labels, and
-		Quanta: The Toll of Unacknowledged Metaphysics, Found. Phys. {\bf
-			21}, 43 (1991).
+		\bibitem{fun} E. Artin: Geometric algebra, Interscience Publishers Inc
+		(1957)
+		\bibitem{sch}E. Schr\"odinger, Annalen der Physik {\bf 102}, 81
+		(1926); English translation in E. Schr\"odinger, {\em Collected
+			papers on Wave Mechanics} (Blackie \& Son, London, 1928).
 		\bibitem{hartle}J.B. Hartle, Quantum Mechanics of Individual Systems,

### 2020-12-01 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
After review with Christine

@@
 Violation of the second requirement would mean that some composite states would not define all the properties of the subsystems.
-That is, while the systems $A$ and $B$ by itself would define an observable $O_A$ and $O_B$, when put grouped together all values would be assigned zero probability; those observables no longer exist.
+That is, while the systems $A$ and $B$ by themselves would define observables $O_A$ and $O_B$, when grouped together all values would be assigned zero probability (XXX); those observables no longer exist.
 In this case, the nature of the system would have changed so radically we would no longer call it a composite system.
@@
 		could equivalently choose one or the other. For
-		example, Schr\"odinger in the papers in which he
-		introduces the Schr\"odinger equation, he writes it
+		example, Schr\"odinger, (XXX) in the papers in which he
+		introduces the Schr\"odinger equation, (XXX) writes it
 		with both signs, as the choice of sign of the imaginary part of the wave function is arbitrary: one sign refers to
@@
 
-	The above proof relies only on the independent preparation of subsystem and not their measurement.
-However, during the review process, an anonymous referee contributed a sketch for a proof that shows very directly that the state and measurement postulates also imply independence of measurements.
+	The above proof relies only on the independent preparation of subsystems and not their measurements.
+(XXX) However, during the review process, an anonymous referee contributed a sketch for a proof that shows very directly that the state and measurement postulates also imply independence of measurements.
 The insight is that the mixture created by all possible measurement outcomes on $B$ must behave overall as a pure state on $A$.
@@
 		Let $\pj{\psi}$ and $\pj{\phi}$ be two preparations for $A$ and $B$.
-Let $\pj{a}$ and $\pj{b}$ two measurements on the respective systems.
+Let $\pj{a}$ and $\pj{b}$ be (XXX) two measurements on the respective systems.
 Then $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi}) P(\pj{b} | \pj{\phi})$.

### 2020-12-01 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex
After review with Christine

@@
 
-	We start from the natural definition of a composite system as the set
+Let us first provide a conceptual overview of our approach. (XXX) We start from the natural definition of a composite system as the set
 of two (or more) quantum systems. The composite system is therefore
@@
 fact that the composite system is fully described by the observables
-	of $A$ and $B$, allows to extend the construction to the composite
+of $A$ and $B$, allows us (XXX) to extend the construction to the composite
 states that do not satisfy statistical independence (the entangled
@@
 state is identified by a ray $\pj{\psi}$, a subspace of the system's
-	Hilbert space spanned by all vectors $\psi$ differing by their
-	(nonzero) modulus and phase: a two-dimensional subspace (a plane). In
+Hilbert space comprising (XXX) all vectors $\psi$ differing by their
+(nonzero) modulus and phase: a one-dimensional complex subspace (a complex plane). In
 the same way, constraining the observable $X$ to a particular outcome
-	value $x_0$ means identifying the subspace spanned by all
+value $x_0$ means identifying the subspace comprising (XXX) all
 non-normalized eigenvectors $|x_0\>$ of arbitrary phase such that
@@
 theorem of projective geometry; (H3)~span surjectivity: the span of
-	the image of map coincides with the full composite Hilbert space.  We
+the image of the map coincides with the full composite Hilbert space.  We
 then prove that, if these three conditions H1, H2 and H3 hold, then
 the map $m$ is the tensor product, namely the Hilbert space of the
-	composite system is a tensor product of the components: the tensor
-	product ``postulate'', which hence loses its status of a postulate. An
+composite system is the (XXX) tensor product of the components' Hilbert spaces (XXX). The tensor
+product ``postulate'' hence loses its status of a postulate. (XXX) An
 overview of all these logical implications is given in
@@
 property $X$, described by the operator with spectral decomposition
-	$\sum_{x,i}x|x_i\>\<x_i|/|\<x_i|x_i\>|$ ($i$ a degeneracy index),
+$\sum_{x,i}x\frac{|x_i\>\<x_i|}{\<x_i|x_i\>}$ ($i$ a degeneracy index),
 returns a value $x$ given that the system is in state $\pj{\psi}$ is
-	$p(x|\psi)=\sum_i|\<\psi|x_i\>|^2/|\<x_i|x_i\>\<\psi|\psi\>|$ (Born
+$p(x|\psi)=\sum_i\frac{|\<\psi|x_i\>|^2}{\<x_i|x_i\>\<\psi|\psi\>}$ (Born
 rule). (c)~The state space of a composite system is given by the
@@
 factors through $T$ uniquely.  This means that there exists a {\em
-		unique} $\hat i$, dependent on $m$, such that $\hat i \circ T=m$.
+  unique} $I$, dependent on $m$, such that $I \circ T=m$.
 In other words, the following diagram commutes:
 \begin{center}
-		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "\hat{i}"] \\
+		\begin{tikzcd}\mathcal{A}\times\mathcal{B} \arrow[rd, "m"]\arrow[r, "T"] & \mathcal{A}\otimes\mathcal{B}\arrow[d, "I"] \\
 			& \mathcal{C}
@@
 the tensor product we can find a unique linear operator
-	$\hat{i} : \mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that
-	$m(a, b) = \hat i(a \otimes b)$. The set $\{ \hat i(a_i\otimes b_j)$
+$I : \mathcal{A} \otimes \mathcal{B} \to \mathcal{C}$ such that
+$m(a, b) = I(a \otimes b)$. The set $\{ I(a_i\otimes b_j)$
 with $|a_i\>$ and $|b_j\>$ orthonormal bases for $\cal A$ and
 ${\cal B}\}$ forms a basis for $\cal C$, since
-	$\hat i(a_i\otimes b_j)=m(a_i,b_j)$ and we have shown above that the
+$I(a_i\otimes b_j)=m(a_i,b_j)$ and we have shown above that the
 latter is a basis.  Thus,
 \begin{align} 
-	&\<\hat i(a_i\otimes b_j)|\hat i(a_k\otimes b_\ell)
+  &\<I(a_i\otimes b_j)|I(a_k\otimes b_\ell)
     \>_{\mathcal{C}}=\<m(a_i,
@@
 $|a_i\otimes b_j\>$ is a basis of the tensor product space
-	${\cal A}\otimes{\cal B}$. Since the function $\hat{i}$ is a linear
+${\cal A}\otimes{\cal B}$. Since the function $I$ is a linear
 function that maps an orthonormal basis of ${\cal A}\otimes{\cal B}$
-	to an orthonormal basis of $\cal C$, $\hat{i}$ is a an isomorphism (a
+to an orthonormal basis of $\cal C$, $I$ is a an isomorphism (a
 bijection that preserves the mathematical structure) between
@@
 ${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert
-	spaces, they are mathematically equivalent: we can loosely say that
-	$\hat i$ is the identity. So we can directly use the
+spaces, they are mathematically equivalent: $c \in {\cal C}$ and $I^{-1}(c)$ (XXX) represent the same physical object.
+In this sense, we can loosely say that $I$ is the identity, as it connects spaces that are physically
+equivalent. So we can directly use the
 tensor product to represent the composite state space. This means that
@@
 $\otimes : \cal A \times \cal B \to \cal A \otimes \cal B$ in the
-	sense that $m \circ \hat i^{-1} = \otimes$.$\square$
+sense that $m \circ I^{-1} = \otimes$.$\square$
 
@@
 $m$ is a bilinear map, we are assured that there exists a unique
-	$\hat i$ such that $\hat i \circ T=m$. Since we show that $\hat i$ is
-	an isomorphism, then $\hat{i}$ bijectively maps vectors in $\cal C$
-	onto vectors in the tensor product space. Namely $m={T}=\otimes$.
+$I$ such that $I \circ T=m$. Since we show that $I$ is
+an isomorphism, then $I$ bijectively maps vectors in $\cal C$
+onto vectors in the tensor product space. Namely $m={T}=\otimes$. (XXX)
 

### 2020-12-01 Articles/2019-QuantumComposition/Submissions/PhysLetters/response.tex
After review with Christine

@@
 paper we had failed to make this distinction clear. (Indeed we had
-wrongly claimed that the supplementary material contained the same
+inaccurately (XXX) claimed that the supplementary material contained the same
 material formulated in a different manner.)
@@
 statistical independence in sufficient detail, and that is actually
-one of the critical step in the demonstration (i.e. making explicit an
+one of the critical steps (XXX) in the demonstration (i.e. making explicit an
 implicit assumption). We rewrote and expanded the part of the paper
@@
 2. I would say that physicists are familiar with the
-notion of 'colinear'.  Hence, I would either (i) provide a definition
-or (ii) express the argument with the more familiar notions 'linear'
-and 'anti-linear' I would be inclined for the second (if possible).
+notion of `colinear'.  Hence, I would either (i) provide a definition
+or (ii) express the argument with the more familiar notions `linear'
+and `anti-linear' I would be inclined for the second (if possible).
 \end{quote}
@@
 \begin{quote}
-3. Regarding 'the dual-space description of a quantum
+3. Regarding `the dual-space description of a quantum
 system through bras is equivalent to the description using kets', what
-means 'equivalent' here? Can you justify it with an argument instead
-of the anecdote that 'in his first papers Schroedinger used both signs
+means `equivalent' here? Can you justify it with an argument instead
+of the anecdote that `in his first papers Schroedinger used both signs
 in his equation, effectively writing two equivalent equations with
@@
 \begin{quote}
-4. Regarding 'The function m, then, is an isomorphism over
+4. Regarding `The function m, then, is an isomorphism over
 all elements of the basis $|a_i\>|b_j\>$'. What it means to be an
-'isomorphism over' some vectors?
+`isomorphism over' some vectors?
 \end{quote}
@@
 elements of the basis into elements of a basis of a tensor product
-space. We understand our notation could lead to confusion. We have
-rewrote that part, and also changed the (admittedly confusing)
+space. We understand our notation could lead to confusion. We rewrote (XXX) that part, and also changed the (admittedly confusing)
 notation to clarify the whole thing.
@@
 \begin{quote}
-5. Regarding 'other linear products of linear spaces
+5. Regarding `other linear products of linear spaces
 exist'. Is the direct product of two vector spaces a vector space? (I
@@
 We have changed that sentence to make it clearer, and have also added
-a footnote to further clarify the distinction
+a footnote to further clarify the distinction. (XXX)
 
@@
 
-We added the following footnote, hoping it will succinctly highlight
-the issue: For example, in the tensor product $a \otimes (b+c) = a
-\otimes b + a \otimes c$ while in the direct product $a \times (b+c) =
-a \times b + 0 \times c$ where $0$ is the zero vector. Also, in the
-tensor product $r * (a \otimes b) = (r*a) \otimes b = a \otimes (r*b)$
-while in the direct product $r * (a \times b) = r*a \times r*b$.
-
-This also addresses the next point.
+We added a footnote, hoping it will succinctly highlight
+the issue. (XXX)
 
 \begin{quote}
-6. Regarding '(Dirac) adds the seemingly innocuous
+6. Regarding `(Dirac) adds the seemingly innocuous
 request'. I understand that requiring the product of vectors to be
@@
 anti-symmetrized tensor product) in those cases. As we explicitly say
-elsewhere, electrons cannot be properly be considered as quantum
+elsewhere, electrons cannot be properly (XXX) considered as quantum
 subsystems (at least in the case in which indistinguishability plays a
@@
 
-We now made this more explicit after the definition: Requirement 1
-ensures that the composite system is well defined at least when the
-components are prepared independently. Conceptually, this ensures that
-the composite system contains all the properties of the
-components. Note that superselection rules or other restrictions may
-prevent the independent preparation of all possible pairs (e.g.~two
-fermions cannot be jointly prepared in the same state). The tensor
-product is recovered only when all pairs are compatible.
+We now made this more explicit after the definition, with a more detailed physical motivation of the requirements. (XXX)
 
@@
 information section will be only read by people who want to understand
-the full details of our argument. This type of readers will not be
+the full details of our argument. This type of reader (XXX) will not be
 scared by the full generality of the argument. Instead in the main
@@
 defining the JOINT probability. We could remove the wedge symbol and
-replace it with the connective "and". We tried, but the formulas
+replace it with the connective ``and". We tried, but the formulas
 become quite heavy, so we believe that this (minimal) concession to
@@
 measurements--together with a set of assumptions reflecting reasonable
-interpretations of the expressions 'composite system' and 'independent
+interpretations of the expressions `composite system' and `independent
 systems'. The essence of the paper, to my mind, lies in the
@@
 supplementary information. We understand that this distinction was not
-clear in the first version of the paper, and take we take
+clear in the first version of the paper, and (XXX) we take
 responsibility for this (we ourselves had mischaracterized our
@@
 keeping in mind that the main paper is written for a more general
-audience and that mostly contains the physics, and the supplementary
+audience and (XXX) mostly contains the physics, and the supplementary
 information only contains the details of the mathematical proofs.
@@
 with two other standard postulates of quantum theory and thereby aim
-to show that there is some redundancy in the standard postulates."
+to show that there is some redundancy in the standard postulates.
 \end{quote}
@@
 is not a necessary postulate, and believe (more on this below) that
-our result is certainly suitable for publication on PRL. Referee A
+our result is certainly suitable for publication in (XXX) PRL. Referee A
 shares this belief.
@@
 Let us first be precise as to what a postulate is.
-If one looks carefully, all postulates of quantum mechanics links something
+If one looks carefully, all postulates of quantum mechanics link (XXX) something
 defined within the physics world (e.g. states, observables, composite system,
@@
 establishing a link between two different definitions (the physical
-and the mathematical) that have no a priori justification. But each
+and the mathematical) that has (XXX) no a priori justification. But each
 element is well defined within its realm. We are not adding new postulates in the sense that we are not adding any
@@
 a composite system, which is ultimately a definition (NOT a postulate) within the physics
-realms, together with the links already established between state space/Hilbert
-space and measurment/Born rule is enough to demonstrate that the state space of the composite
+realm, (XXX) together with the links already established between state space/Hilbert
+space and measurement/Born (XXX) rule is enough to demonstrate that the state space of the composite
 system must be the tensor product of the subsystems.
@@
 Hilbert space" and something like ``the state of a composite system
-will fully determine the probability distribution of the observable of
+will fully determine the probability distribution of the observables (XXX) of
 all its parts and only of its parts." The first connects two
@@
 a priori reason for that and we are not giving one in this paper. The
-second is defining an object based on the other. That is, we do not
+second is defining one object based on others (XXX). That is, we do not
 define composite systems in some other way and than later we
@@
 
-That "world" would simply correspond to a different definition of a composite system. Whereas we used the natural one: the (set) union of the subsystems. 
+That ``world" would simply correspond to a different definition of a composite system. Whereas we used the natural one: the (set) union of the subsystems. 
 In the standard definition, the
@@
 rather than at the end of that sentence, because we did not want to
-ruin the flow of the narration by introducing the concept of "states"
+ruin the flow of the narration by introducing the concept of ``states"
 in that sentence (which refers to systems).
@@
 \begin{quote}
-2. On page 2, postulate (a) begins, 'The state of a system
+2. On page 2, postulate (a) begins, `The state of a system
 is described by a ray...'. As expressed here, this postulate actually
@@
 each component system is not described by a ray in that system's
-Hilbert space. I'm not sure why the authors didn't say, 'A pure state
+Hilbert space. I'm not sure why the authors didn't say, `A pure state
 of a system is described by a ray...'. (The word 'pure' is missing
@@
 
-Definitely: the word 'pure' is missing! We apologize for this
+Definitely: the word `pure' is missing! We apologize for this
 oversight and thank the Referee for pointing it out. We certainly
@@
 
-In conclusion, we thank the Referee B for his/her insightful
+In conclusion, we thank (XXX) Referee B for his/her insightful
 observations and helpful comments. We believe that his/her assessment
@@
 
-In closing, we were very pleasurably surprised by the quality of both
-reports and we thank both Referees for their time they spent in
+In closing, we were very pleasantly (XXX) surprised by the quality of both
+reports and we thank both Referees for the (XXX) time they spent in
 evaluating our paper.

### 2020-12-05 Articles/2019-QuantumComposition/Submissions/PhysLetters/response.tex

@@
 paper we had failed to make this distinction clear. (Indeed we had
-inaccurately (XXX) claimed that the supplementary material contained the same
+inaccurately claimed that the supplementary material contained the same
 material formulated in a different manner.)
@@
 statistical independence in sufficient detail, and that is actually
-one of the critical steps (XXX) in the demonstration (i.e. making explicit an
+one of the critical steps in the demonstration (i.e. making explicit an
 implicit assumption). We rewrote and expanded the part of the paper
@@
 elements of the basis into elements of a basis of a tensor product
-space. We understand our notation could lead to confusion. We rewrote (XXX) that part, and also changed the (admittedly confusing)
+space. We understand our notation could lead to confusion. We rewrote that part, and also changed the (admittedly confusing)
 notation to clarify the whole thing.
@@
 We have changed that sentence to make it clearer, and have also added
-a footnote to further clarify the distinction. (XXX)
+a footnote to further clarify the distinction.
 
@@
 
-We added a footnote, hoping it will succinctly highlight
-the issue. (XXX)
+We added a footnote which succinctly highlights
+the issue.
 
@@
 where the state space of composite systems is obtained through the
-direct product, and (a,b+c) is DIFFERENT from (a,b)+(a,c).
+direct product, and (a,b+c) is DIFFERENT from (a,b)+(a,c). We explained it better in the new version of the paper.
 
@@
 anti-symmetrized tensor product) in those cases. As we explicitly say
-elsewhere, electrons cannot be properly (XXX) considered as quantum
+elsewhere, electrons cannot be properly considered as quantum
 subsystems (at least in the case in which indistinguishability plays a
@@
 
-We now made this more explicit after the definition, with a more detailed physical motivation of the requirements. (XXX)
+We now made this more explicit after the definition, with a more detailed physical motivation of the requirements.
 
@@
 information section will be only read by people who want to understand
-the full details of our argument. This type of reader (XXX) will not be
+the full details of our argument. This type of reader will not be
 scared by the full generality of the argument. Instead in the main
@@
 
-We agree that we haven't clearly explained the formalism and we have
+We agree that we hadn't clearly explained the formalism and we have
 now done so. We do not really need the formalism of logical
@@
 defining the JOINT probability. We could remove the wedge symbol and
-replace it with the connective ``and". We tried, but the formulas
+replace it with the connective ``AND". We tried, but the formulas
 become quite heavy, so we believe that this (minimal) concession to
@@
 supplementary information. We understand that this distinction was not
-clear in the first version of the paper, and (XXX) we take
+clear in the first version of the paper, and we take
 responsibility for this (we ourselves had mischaracterized our
@@
 keeping in mind that the main paper is written for a more general
-audience and (XXX) mostly contains the physics, and the supplementary
+audience and mostly contains the physics, and the supplementary
 information only contains the details of the mathematical proofs.
@@
 is not a necessary postulate, and believe (more on this below) that
-our result is certainly suitable for publication in (XXX) PRL. Referee A
+our result is certainly suitable for publication in PRL. Referee A
 shares this belief.
@@
 Let us first be precise as to what a postulate is.
-If one looks carefully, all postulates of quantum mechanics link (XXX) something
+If one looks carefully, all postulates of quantum mechanics link something
 defined within the physics world (e.g. states, observables, composite system,
@@
 establishing a link between two different definitions (the physical
-and the mathematical) that has (XXX) no a priori justification. But each
+and the mathematical) that has no a priori justification. But each
 element is well defined within its realm. We are not adding new postulates in the sense that we are not adding any
@@
 a composite system, which is ultimately a definition (NOT a postulate) within the physics
-realm, (XXX) together with the links already established between state space/Hilbert
-space and measurement/Born (XXX) rule is enough to demonstrate that the state space of the composite
+realm, together with the links already established between state space/Hilbert
+space and measurement/Born rule is enough to demonstrate that the state space of the composite
 system must be the tensor product of the subsystems.
@@
 Hilbert space" and something like ``the state of a composite system
-will fully determine the probability distribution of the observables (XXX) of
+will fully determine the probability distribution of the observables of
 all its parts and only of its parts." The first connects two
@@
 a priori reason for that and we are not giving one in this paper. The
-second is defining one object based on others (XXX). That is, we do not
+second is defining one object based on others. That is, we do not
 define composite systems in some other way and than later we
@@
 definition that comes from our physical intuition.
+
+In any case, the Referee's observations were very useful to us: we realized that the connection between postulates and physics should have been explained much better and we did so in the new version of the paper.
 
@@
 
-We understand that our paper was definitely unclear on this and we have
-clarified this point.
+Again, the Referee's comment made us understand that our paper was definitely unclear on this and we have
+clarified it in the new version.
 
@@
 
-In conclusion, we thank (XXX) Referee B for his/her insightful
+In conclusion, we thank Referee B for his/her insightful
 observations and helpful comments. We believe that his/her assessment
@@
 
-In closing, we were very pleasantly (XXX) surprised by the quality of both
-reports and we thank both Referees for the (XXX) time they spent in
+In closing, we were very pleasantly surprised by the quality of both
+reports and we thank both Referees for the time they spent in
 evaluating our paper.

### 2020-12-05 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex

--- 2020-11-25 Articles/2019-QuantumComposition/Submissions/PhysLetters/supplementaryinfo.tex
@@
 Violation of the second requirement would mean that some composite states would not define all the properties of the subsystems.
-That is, while the systems $A$ and $B$ by itself would define an observable $O_A$ and $O_B$, when put grouped together all values would be assigned zero probability; those observables no longer exist.
+That is, while the systems $A$ and $B$ by themselves would define observables $O_A$ and $O_B$, when grouped together all values would be assigned zero probability; those observables no longer exist.
 In this case, the nature of the system would have changed so radically we would no longer call it a composite system.
@@
 		could equivalently choose one or the other. For
-		example, Schr\"odinger in the papers in which he
-		introduces the Schr\"odinger equation, he writes it
+		example, Schr\"odinger, in the papers in which he
+		introduces the Schr\"odinger equation, writes it
 		with both signs, as the choice of sign of the imaginary part of the wave function is arbitrary: one sign refers to
@@
 
-	The above proof relies only on the independent preparation of subsystem and not their measurement.
+	The above proof relies only on the independent preparation of subsystems and not their measurements.
 However, during the review process, an anonymous referee contributed a sketch for a proof that shows very directly that the state and measurement postulates also imply independence of measurements.
@@
 		Let $\pj{\psi}$ and $\pj{\phi}$ be two preparations for $A$ and $B$.
-Let $\pj{a}$ and $\pj{b}$ two measurements on the respective systems.
+Let $\pj{a}$ and $\pj{b}$ be two measurements on the respective systems.
 Then $P(\pj{a} \wedge \pj{b} | \pj{\psi} \wedge \pj{\phi}) = P(\pj{a} | \pj{\psi}) P(\pj{b} | \pj{\phi})$.

### 2020-12-05 Articles/2019-QuantumComposition/Submissions/PhysLetters/tensor.tex

@@
 
-Let us first provide a conceptual overview of our approach. (XXX) We start from the natural definition of a composite system as the set
+Let us first provide a conceptual overview of our approach. We start from the natural definition of a composite system as the set
 of two (or more) quantum systems. The composite system is therefore
@@
 fact that the composite system is fully described by the observables
-of $A$ and $B$, allows us (XXX) to extend the construction to the composite
+of $A$ and $B$, allows us to extend the construction to the composite
 states that do not satisfy statistical independence (the entangled
@@
 state is identified by a ray $\pj{\psi}$, a subspace of the system's
-Hilbert space comprising (XXX) all vectors $\psi$ differing by their
+Hilbert space comprising all vectors $\psi$ differing by their
 (nonzero) modulus and phase: a one-dimensional complex subspace (a complex plane). In
 the same way, constraining the observable $X$ to a particular outcome
-value $x_0$ means identifying the subspace comprising (XXX) all
+value $x_0$ means identifying the subspace comprising all
 non-normalized eigenvectors $|x_0\>$ of arbitrary phase such that
@@
 the map $m$ is the tensor product, namely the Hilbert space of the
-composite system is the (XXX) tensor product of the components' Hilbert spaces (XXX). The tensor
-product ``postulate'' hence loses its status of a postulate. (XXX) An
+composite system is the tensor product of the components' Hilbert spaces. The tensor
+product ``postulate'' hence loses its status of a postulate. An
 overview of all these logical implications is given in
@@
 ${\cal C}\cong{\cal A}\otimes{\cal B}$ are isomorphic as Hilbert
-spaces, they are mathematically equivalent: $c \in {\cal C}$ and $I^{-1}(c)$ (XXX) represent the same physical object.
+spaces, they are mathematically equivalent: $c \in {\cal C}$ and $I^{-1}(c)$ represent the same physical object.
 In this sense, we can loosely say that $I$ is the identity, as it connects spaces that are physically
@@
 an isomorphism, then $I$ bijectively maps vectors in $\cal C$
-onto vectors in the tensor product space. Namely $m={T}=\otimes$. (XXX)
+onto vectors in the tensor product space. Namely $m={T}=\otimes$.
 

### 2020-12-07 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Edits to section 2 on defining classical system and section 6.3 on Curiel's counterarguments

@@
 
+Like North \parencites*[85]{North}, our argument involves investigating Hamiltonian systems outside the hyper-regular domain.
+
 Thesis: despite the mathematical equivalence of Hamiltonian and Lagrangian mechanics within the hyper-regular domain, the two frameworks have important differences.
@@
 
-Provide any relevant background on classical mechanics, specifically to clarify/define terminology.
+Goal of section: Provide any relevant background on classical mechanics, specifically to clarify/define terminology.
 Might be able to just put this in the introduction proper, or the start of the next section.
@@
 
-A classical Hamiltonian system is any system whose phase space is a symplectic manifold and that evolves according to Hamiltonian evolution (i.e. according to a symplectomorphism).
+% [To do: figure out if North and/or Curiel restrict their discussion/argument to the hyper-regular domain].
+It seems as though Jill North is explicitly not assuming that her argument applies only to the hyper-regular domain [page 85, especially footnote 50].
+Part of her argument involves motivating the idea that we will not in general be able to construct an equivalent Lagrangian formulation, and this involves appealing to systems outside the hyper-regular domain.
+% Figure out explicitly how North defines classical systems.
+It might be in terms of the equations of motion that hold for classical particles [87].
+North is arguing that ``symplectic structure is a simpler, more natural, more general underlying structure" that ``suffices for" this theory of classical mechanics [87].
+Seems similar to GC position.
 
-Throughout, we will focus only on non-dissipative classical systems, i.e.
-conservative systems.
-For in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $.
-This requires that the system is conservative, which is equivalent to the system being deterministic and reversible.
+Various classes of classical systems are commonly delimited by specifying a set of equations that they satisfy, such as Hamilton's equations or the Euler--Lagrange equations.
+We will instead delimit classical systems based on the physical conditions that they satisfy, such as reducibility and deterministic evolution.
+We will then derive the relevant equations as consequences of these physical assumptions.
+For instance, we will see that a classical Hamiltonian system is any system whose phase space is a symplectic manifold and that evolves according to Hamiltonian evolution (i.e.
+according to a symplectomorphism).
+
+The assumption of reducibility plays the key role of distinguishing classical systems from quantum-mechanical ones.
+Unlike classical systems, quantum systems are \textit{irreducible}, in the sense that specifying the state of the whole system is not equivalent to specifying the states of its parts (and vice versa).
+In contrast, a reducible system comprises subsystems whose states jointly determine the state of the whole system.
+In the widest sense, a classical system is a reducible one.
+
+Newtonian systems do not necessarily have deterministic and reversible evolution, for instance in the case of a damped harmonic oscillator.
+Yet, they do satisfy the motion equivalence assumption because knowledge of the position and velocity allows for the determination of the acceleration [why does this entail that motion equivalence is satisfied?].
+In contrast, Hamiltonian systems are necessarily deterministic and reversible, although they need not satisfy motion equivalence.
+We will understand Lagrangian mechanics as the intersection of Newtonian and Hamiltonian systems: they have deterministic and reversible evolution, while also satisfying motion equivalence.
+
+By focusing on systems with deterministic and reversible evolution, we necessarily exclude dissipative classical systems (i.e.
+ones that do not conserve energy).
+For determinism and reversibility are equivalent to a system being conservative.
+This restriction arises partly from the following reason: in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we will assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $.
+This requires that the system is conservative.
 Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
-% Dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential
+We, of course, do not deny the sense in which the Lagrangian formulation applies to dissipative systems, but this is not what we will mean by ``Lagrangian mechanics" \parencites[]{Smith}{Curiel}.\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems (by moving to the extended phase space), although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime.
+We will view dissipative systems as Newtonian, but neither Lagrangian nor Hamiltonian systems [appeal to Norton's dome for motivation?]
+
+Further motivation for focusing on non-dissipative systems: dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential.
+Or, as Curiel puts it, the systems involve a ``force $1 $-form that is not exact and that is not Lie-derived along the Lagrangian vector field solving the inhomogeneous system" [311].
+
+% Question of how Newtonian mechanics fits and: we can understand Newtonian mechanics as a hybrid between Hamiltonian and Lagrangian mechanics.
+On the one hand, Newtonian systems need not be deterministic and reversible.
+On the other hand, they satisfy the motion equivalence assumption, because knowledge of the position and velocity allows for the determination of the acceleration.
 
@@
 
+\url{https://physics.stackexchange.com/questions/398159/formulation-of-an-action-principle-in-a-general-phase-space-no-cotangent-bundle?rq=1}
+
+seems to have very similar kind of argument
+
 % Somewhere below, emphasize that you have a unique trajectory only if the action is a minimum or maximum (if it is just stationary but not a minimum or maximum, then you do not have a unique trajectory)
@@
 Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2)" (306).
+
+Like our argument, part of Curiel's argument involves appealing to systems outside of the hyper-regular domain.
+Curiel acknowledges both that only hyper-regular Lagrangians have unique solutions and that Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}.
+Nevertheless, he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage.
+On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}.
+In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy a particular set of physically significant kinematical constraints (his equations 4.3).
+Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will still satisfy the kinematical constraints of Lagrangian mechanics \parencites*[311]{Curiel}.
+
+Summary: Hamiltonian mechanics provide solutions outside hyper-regular domain (but not these kinematical constraints) whereas Lagrangian mechanics preserves these kinematical constraints outside hyper-regular domain.
+Worry: perhaps GC and Curiel are in a verbal dispute about the proper/privilege definition of a classical system (with Curiel privileging these kinematical constraints, and GC privileging deterministic/reversible evolution).
+
+On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems).
+In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers Hamiltonian $H=p$, which corresponds to system moving with constant velocity (e.g.
+velocity of 1).
+So this could very well be a model of a photon, moving at the speed of light.
+Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating Euler--Lagrange equation.
+Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310].
+But then does Curiel consider this system to not be a classical system?
+
+Question to ask GC: does he agree with the following claims made by Curiel regarding the differences between the symplectic structure in Hamiltonian mechanics (which is fixed) vs.
+the symplectic structure that arises from the Euler--Lagrange equations (which varies from Lagrangian to Lagrangian): \\
+Hamiltonian mechanics: ``in Hamiltonian mechanics, the fact that the symplectic structure is fixed once and for all enforces the satisfaction of the kinematical constraints (5.3), and it encodes the existence and uniqueness of solutions to the equations of motion and the conservation of energy, among many other things.
+All of these features are independent of the individual Hamiltonian used in the equation" [310].
+\\
+Lagrangian mechanics: ``In Lagrangian mechanics, in contrast, it is the Lagrangian itself that encodes existence and uniqueness of solutions (which will obtain when and only when the Lagrangian is regular), and there is not even a guarantee that energy is conserved, as one can, for example, represent dissipative systems in the Lagrangian framework" [310].
+
+% Either way, presumably we can agree with Curiel's explanation/understanding for why the Lagrangian must be regular: Curiel arguing that he can give an account of ``why it makes physical sense to demand that the Lagrangian be regular", i.e.
+satisfy equation 4.6 for the Hessian determinant.
+He notes that by considering the variational problem, we understand ``why the Lagrangian must be quadratic (at least) in the velocities for there to be a unique solution to the Euler--Lagrange equation" [308].
+Argument: for the variational problem to be well-posed, you need to be able to ``take non-trivial derivatives of the Lagrangian up to second-order in the velocity, since the crucial integration by parts that yields the Euler--Lagrange equation demands this, on pain of giving, for example, the tautology $0=0$" in some cases, or a contradiction in others.
 

### 2020-12-14 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Updated sections 1, 2, 5, and 6
Rewrote large parts and left notes for remaining to do's.

--- 2020-11-30 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 \begin{document}
-%% testing from desktop
 
-\title{TBD}
+\title{From physical conditions to Hamiltonian and Lagrangian mechanics}
 \author{TBD}
@@
 
-Current section ideas:
-
-\begin{enumerate}[1.)]
-
-\item Assumptions of physics framework section
-
-\begin{itemize}
-
-\item motivate everything upfront: going to show that Hamiltonian mechanics corresponds to two conditions being satisfied, while Lagrangian mechanics requires a third condition
-\item define the two assumptions, and give motivation
-\item figuring out which mathematical structures are required for systems that satisfy certain physical conditions
-\item go over the upshot of the philosophical argument, specifically how this avoids the worries about measuring the amount of structure that plagued North (i.e. Swanson and Halvorson paper)
-\item Then, give the derivation, as gentle and well-motivated as possible. But would be ideal if readers could skip or skim this section and still get the philosophical upshot, relatively early on.
-
-\end{itemize}
-
-\item Physical significance of invariant distributions
-
-\begin{itemize}
-
-\item explain how these provide a novel way of understanding things, one that we can physically privilege
-
-\item connect this with North on invariant structure, but disavow metaphysical commitments
-
-\end{itemize}
-
-\item How to interpret Lagrangian mechanics in light of Hamiltonian mechanics
-
-\begin{itemize}
-
-\item gauge transformation stuff
-\item go over the cool stationary action principle derivation/argument
-\item clarify the status of Riemannian metric structure in the Lagrangian formalism
-\item respond to Curiel's privileging of Lagrangian mechanics
-\item respond to Curiel's argument that Hamiltonian mechanics involves the imposition of an ad hoc condition in order to get the right kinematical constraints, whereas these fall out naturally from Lagrangian mechanics.
-
-\end{itemize}
-
-\item Considerations in light of categorical equivalence proofs
-
-\begin{itemize}
-
-\item  clarify that argument is not in tension with categorical equivalence within the hyper regular domain. make this precise.
-\item Respond to the worry that something must have gone amiss if we have a difference in the number of conditions required for the Hamiltonian vs.
-Lagrangian mechanics, yet still have categorical equivalence in the hyper-regular domain
-
-\end{itemize}
-
-\end{enumerate}
-
 \section{Introduction}
+\label{introduction}
 
@@
 Categorical equivalence provides a natural interpretive criterion for theoretical equivalence because it yields mutual inter-translatability between theory formulations.
-At first glance, this equivalence seems to motivate the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain.
-This equivalence undermines prior positions that have privileged either Hamiltonian mechanics or Lagrangian mechanics on physical or metaphysical grounds \parencites[]{North}{Curiel}.
+At first glance, this equivalence motivates the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain.
+Yet, it allows for physical and conceptual differences to arise outside the hyper-regular domain.
+%Once identified, these differences can even affect how we interpret hyper-regular systems.
+
+Within this wider domain, \textcites[]{North} has argued for privileging Hamiltonian mechanics, on the grounds that it requires less mathematical structure to formulate.
+In contrast, \textcites[]{Curiel} recommends privileging Lagrangian mechanics, arguing that it isolates physically-significant kinematical constraints that all classical systems must satisfy.
 Here, we will argue that there are indeed physical grounds for privileging Hamiltonian mechanics over Lagrangian mechanics, their categorical equivalence notwithstanding.
-Compared to Hamiltonian systems, an additional physical condition must be satisfied for a classical system to be Lagrangian.
+Whereas North's argument relies on a problematic criterion for comparing structure---based on the size of symmetry groups \parencites[]{Swanson}---our argument avoids this pitfall.
+We show that a system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian.
+To be Lagrangian, systems must satisfy an additional physical condition, making Lagrangian mechanics a proper subset of Hamiltonian mechanics.
 Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this difference has interpretive consequences even within the hyper-regular domain.
@@
 
-Thesis: despite the mathematical equivalence of Hamiltonian and Lagrangian mechanics within the hyper-regular domain, the two frameworks have important differences.
-In particular, Hamiltonian systems arise when a smaller number of basic physical conditions are met.
-A Lagrangian system must satisfy a further physical condition.
-This shows that Hamiltonian mechanics is in a sense more fundamental.
-Defending this claim does not require commitments to fundamental structure or perfectly natural properties.
-(Furthermore, the Hamiltonian framework is more general because a Hamiltonian defines a unique solution (unique evolution in phase space) even when the Hessian is zero.
-In contrast, whenever the Lagrangian is not hyper-regular, the Euler--Lagrange equations do not possess a unique solution.
-Additionally, density plays a special role in the Hamiltonian framework, whereas it is in some sense an accident that the density in configuration space is invariant (simply because the Jacobian equals unity in Cartesian coordinates).
+Understanding the relationship between Hamiltonian and Lagrangian mechanics depends partly on how one defines ``classical systems." Different definitions can result in different interpretive relationships.
+For instance, restricting classical systems to the hyper-regular domain precludes certain interpretive considerations, To avoid verbal disputes, it is essential to clarify what falls within the scope of ``classical systems." One way to delimit classical systems is to specify a set of equations that they satisfy---such as Hamilton's equations or the Euler--Lagrange equations.
+We will instead follow a different approach by delimiting classical systems based on the physical conditions that they satisfy.
+Section~\ref{infinitesimal} introduces the condition of \textit{infinitesimal reducibility}, meaning that the state of the whole system is equivalent to specifying the states of its parts (and vice versa).
+Infinitesimal reducibility distinguishes classical from quantum systems, where the latter are irreducible (we cannot specify the whole state by specifying the states of its subsystems).
+In the widest sense, a classical system is any infinitesimally reducible system.
+%a reducible system comprises subsystems whose states jointly determine the state of the whole system.
+
+Section~\ref{deterministic} introduces the joint condition of deterministic and reversible evolution.
+We will show that a classical system which satisfies this condition is Hamiltonian: its phase space is a symplectic manifold that evolves according to Hamiltonian evolution  (i.e.
+according to a symplectomorphism).
+Following a deterministic and reversible evolution is equivalent to being a conservative system, i.e.
+one that conserves energy.
+Hence, our discussion of Hamiltonian mechanics necessarily excludes dissipative systems, i.e.
+ones that do not conserve energy.
+Likewise, our derivation of Lagrangian mechanics in Section~\ref{motion} will also take place within the context of conservative systems.
+Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
+We, of course, do not deny the sense in which the Lagrangian formulation applies to dissipative systems, but this is not what we will mean by ``Lagrangian mechanics" \parencites[]{Smith}{Curiel}.\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime.
+We will view dissipative systems as Newtonian, but neither Lagrangian nor Hamiltonian.
+
+%Further motivation for focusing on non-dissipative systems: dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential.
+Or, as Curiel puts it, the systems involve a ``force $1 $-form that is not exact and that is not Lie-derived along the Lagrangian vector field solving the inhomogeneous system" [311].
+
+We begin in Section~\ref{privileging} with an informal overview of our argument for privileging Hamiltonian mechanics.
+We introduce the key physical conditions that distinguish Hamiltonian from Lagrangian mechanics.
+Section~\ref{fundamentality} explains how our position differs from North's.
+Specifically, our argument does not require commitments to fundamental structure or perfectly natural properties.
+Section~\ref{derivation} describes in detail the derivation of Hamiltonian mechanics for systems that satisfy two physical conditions: infinitesimal reducibility and deterministic/reversible evolution.
+We then show (Section~\ref{motion}) that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
+Section~\ref{density} explains how the invariant phase space density plays a special role in the Hamiltonian framework.
+In contrast, we show that in configuration space, the invariance of the density is coordinate-dependent (arising from the fact that the Jacobian equals unity in Cartesian coordinates but not in general).
+Finally, Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework.
+We provide a geometric interpretation of the action principle, which applies even outside the hyper-regular domain.
+We also consider and rebut objections stemming from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
+% In configuration space, it is in some sense an accident that the density is invariant
 
@@
 
-Targets: Curiel's privileging of Lagrangian mechanics and argument that Hamiltonian mechanics is ad hoc; North's somewhat flawed privileging of Hamiltonian mechanics on the basis of it requiring less structure; a motivation stemming from categorical equivalence proofs for thinking that there are no physically significant differences between the Hamiltonian and Lagrangian frameworks (at least within the hyper regular domain).
+%It seems as though North is explicitly not assuming that her argument applies only to the hyper-regular domain [page 85, especially footnote 50].
+Part of her argument involves motivating the idea that we will not in general be able to construct an equivalent Lagrangian formulation, and this involves appealing to systems outside the hyper-regular domain.
+% To-do: Figure out how North defines classical systems.
+It might be in terms of the equations of motion that hold for classical particles [87].
+North is arguing that ``symplectic structure is a simpler, more natural, more general underlying structure" that ``suffices for" this theory of classical mechanics [87].
+Seems similar to GC position.
 
-Goals: provide a gentle introduction to Assumptions of Physics framework, and show that it has real philosophical payoffs for interpreting physical theories, such as classical mechanics.
-
-\section{Defining `classical mechanical systems'}
-
-Provide any relevant background on classical mechanics, specifically to clarify/define terminology.
-Might be able to just put this in the introduction proper, or the start of the next section.
-Make explicit how our terminology relates to other standard terminology in the philosophical literature.
-
-A classical Hamiltonian system is any system whose phase space is a symplectic manifold and that evolves according to Hamiltonian evolution (i.e. according to a symplectomorphism).
-
-Throughout, we will focus only on non-dissipative classical systems, i.e.
-conservative systems.
-For in the context of transforming between the Hamiltonian and Lagrangian formulations (via a Legendre transform), we assume that the conjugate momentum $p$ is the derivative of the Lagrangian with respect to the generalized velocity $v $.
-This requires that the system is conservative, which is equivalent to the system being deterministic and reversible.
-Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
-% Dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential
+% In the introduction, should we give glosses on hyper-regular, infinitesimally reducible, deterministic/reversible, and motion equivalence? Otherwise, discussion might be less accessible.
 
 \section{Privileging Hamiltonian mechanics}
+\label{privileging}
 
-Our framework develops a correspondence between three physical conditions and their associated mathematical formalisms.
-In this way, we straightforwardly interpret the physical significance of these mathematical formalisms.
-First, a system is \textit{infinitesimally reducible} provided that the states of its constituent, infinitesimal particles completely specify the overall state of the system.
+%[To do: Do we agree with the following claim defended by North: ``as far as we can tell, we need only symplectic structure to do classical mechanics" [74].
+If so, is our position in tension with the more general formulation of Lagrangian mechanics noted by Swanson and Halvorson, in terms of the almost tangent structure of the tangent bundle?
+[See their footnote 9, pages 4-5]
+
+In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by ``fundamental." Of the various notions of fundamentality, we will mean something entirely non-mysterious: a formalism is more fundamental provided that it is logically more general.
+In this case, the Lagrangian framework is less general because it requires that an additional condition hold, compared to the Hamiltonian framework.
+Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.
+
+We now outline the three main steps of our argument, summarized in Figure~\ref{diagram}.
+The left column lists three basic physical conditions: infinitesimal reducibility, determinism/reversibility, and motion equivalence.
+Each condition entails physically significant properties, described in the middle column.
+Each physical property corresponds to a formal mathematical framework depicted in the right column: including phase space, Hamiltonian mechanics, and Lagrangian mechanics.
+In this way, our framework develops a correspondence between physical conditions and their associated mathematical formalisms.
+We thereby gain a straightforward interpretation of the physical significance of these formalisms.
+%Of the various notions of fundamentality, there is at least the following non-mysterious definition:
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{Diagram.png}
+\caption{Physical conditions and corresponding properties and formalisms}
+\label{diagram}
+\end{figure}
+
+First, a system is \textit{infinitesimally reducible} provided that the states of its constituent, infinitesimal parts completely specify the overall state of the system.
+Colloquially, we can say that the whole state is equal to the sum of its parts, where we take these infinitesimal parts to be classical particles.
 When this condition is satisfied, we can define a distribution $\rho$ over particle states.
@@
 The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
+[To do: Clarify the scope of our treatment: is it restricted to $N$-particle classical mechanics?
+Or does it also handle rigid bodies and classical continual mechanics?
+Cite Mark Wilson 2013 on these issues].
+% % I think I could spend more time clarifying reducibility, infinitesimal. Maybe define separately first?
 
 Secondly, a system has \textit{deterministic and reversible evolution} provided that the present state of the system uniquely determines both its future and past states.
-This condition entails that the preceding invariant distribution is in fact a conserved phase space density.
+This condition entails that the invariant distribution $\rho$ is in fact a conserved phase space density.
 We can thus model such systems using Hamiltonian mechanics, defined on a phase space with a symplectomorphism.
 These two conditions thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution.
-Section~\ref{deterministic} provides this derivation in detail.
+Section~\ref{deterministic} justifies this correspondence.
+[To do: on the topic of failures of determinism, cite Earman  \textit{a primer on determinism}, Baez singularities paper, or Norton's dome paper].
 
-Finally, a system satisfies a further condition---motion equivalence---when its trajectory in physical space determines and is determined by its evolution in phase space.
-This condition leads to a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics in the hyper-regular domain (Section~\ref{motion}).
-Hence, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
-To see this, note that some systems fail to satisfy the condition of motion equivalence, i.e.
-it is not always true that we can determine the evolution of a system in phase space from its trajectory in physical space.
-For instance, knowledge of the trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from its trajectory alone.
+Finally, some classical systems satisfy a further condition, which we call \textit{motion equivalence}.
+Motion equivalence holds when the system's trajectory in physical space determines---and is determined by---its evolution in phase space.
+In other words, knowledge of the trajectory in physical space (e.g.
+Euclidean three-space $\mathbb{R}^3$) suffices for knowledge of the trajectory in phase space, and vice versa.
+Physically, knowledge of the trajectory in physical space corresponds to knowledge of position and velocity, while knowledge of the trajectory in phase space requires knowing the system's momentum.
+Hence, motion equivalence entails a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics in the hyper-regular domain (Section~\ref{motion}).
+As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
+
+Importantly, some systems fail to satisfy the condition of motion equivalence.
+It is not always true that we can determine the evolution of a system in phase space from its trajectory in physical space.
+This is because knowing a system's position and velocity is not in general sufficient for knowing its momentum.
+For instance, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from only its trajectory through space.
 In general, different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
-In contrast, the trajectories in phase space (encoded by the Hamiltonian vector field) uniquely define the Hamiltonian up to a constant.} Hence, the set of classical systems that satisfy this further physical condition is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics.
-On these grounds, we will argue in Section~\ref{fundamentality} that Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
-[To do: verify that the scenario of two different Lagrangians giving rise to the same trajectories in physical space with different energies, arises only when motion equivalence condition fails.
+In contrast, the trajectories in phase space (encoded by the Hamiltonian vector field) uniquely define the Hamiltonian up to a constant.} Hence, the set of classical systems that satisfy motion equivalence is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics.
+For at least this reason, Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
+[To do: verify that the scenario of two different Lagrangians---giving rise to the same trajectories in physical space with different energies---arises only when the motion equivalence condition fails.
 See Curiel 2014, page 42 for his take on this kind of scenario.]
+
+%Figure~\ref{diagram} summarizes our characterization of Hamiltonian and Lagrangian mechanics, in terms of physical conditions and corresponding mathematical structures.
+Section~\ref{derivation} justifies the logical connection between these physical conditions and corresponding mathematics.
+Before delving into this derivation, Section~\ref{fundamentality} discusses its philosophical significance.
 
@@
 
-Figure~\ref{diagram} summarizes our characterization of Hamiltonian and Lagrangian mechanics, in terms of physical conditions and corresponding mathematical structures.
-The left column lists the three physical conditions: infinitesimal reducibility, determinism/reversibility, and motion equivalence.
-Each condition entails physically significant properties, described in the middle column.
-Each of these physical properties corresponds to a formal mathematical framework, described in the right column.
-Section~\ref{derivation} justifies the logical connection between these physical conditions and corresponding mathematics.
-Before delving into this derivation, Section~\ref{fundamentality} discusses its philosophical significance.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\textwidth]{Diagram.png}
-\caption{Physical conditions and corresponding properties and formalisms}
-\label{diagram}
-\end{figure}
-
-\subsection{Fundamentality and structure}
+\subsection{(Non-metaphysical) fundamentality and structure}
 \label{fundamentality}
 
-In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by ``fundamental." Of the various notions of fundamentality, there is at least the following non-mysterious definition: a formalism is more fundamental provided that it is logically more general.
-In this case, the Lagrangian framework is less general because it requires that an additional condition hold, compared to the Hamiltonian framework.
-Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.
-In particular, the assumption of motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.
-This notion of fundamentality leads as well to a concomitant notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied.
-In Section~\ref{distribution}, we will discuss another sense in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.
+% North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory.
+Structure comprises those quantities that remain intact when we transform one allowable coordinatization of the laws into another" [78]
+
+% North's epistemic method for figuring out the structure: ``look at what is invariant under allowable transformations"; this tells us what the structure is [73].
+
+As indicated at the start of Section~\ref{privileging}, we use ``fundamentality" in a non-metaphysical fashion.
+On our usage, differences in fundamentality amount to differences in generality.\footnote{Of course, we allow for other non-metaphysical uses of ``fundamental," tracking differences in epistemic or conceptual primacy.
+For instance, we can think of the canonical position and momentum observables as ``epistemically fundamental" because they suffice to completely specify a classical state, fixing its values for all other observables.
+\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{density},discusses another sense in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied.
+Compared with North's \parencites*[]{North} account of structure, our definition has two distinct advantages.
+First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which empiricists and pragmatists decry.
+Secondly, it avoids a technical objection that Swanson and Halvorson \parencites*[]{Swanson} levy against North's more precise mathematical criterion for comparing structure.
+
+Methodologically (but not metaphysically), our approach has many parallels to North's.
+We agree with North that it is fruitful to identify both invariants and---more generally---``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}.
+Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North}.
+Indeed, we undertake precisely this sort of investigation in Section~\ref{derivation}.
+However, North advocates a metaphysical account of structure, about which we remain agnostic.
+According to North, ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" \parencites*[66]{North}.
+These notions of fundamentality and intrinsicality correspond to differences in joint-carving, with more fundamental or intrinsic features being more joint-carving, or in Lewis's \parencites*[]{Lewis1983} terminology, more natural.
+
+Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.
+While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus structure.
+Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration.
+As we will see, precisely this scenario occurs in the context of Lagrangian mechanics.
+Hence, whereas North views Lagrangian mechanics as containing ``excess, superfluous structure" \parencites*[75]{North}, we take this additional structure as unnecessary but still physically significant.
+It corresponds to the physical assumption of motion equivalence, which applies to many classical systems.
+%This characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system. 
+% classical mechanics also deploys physically significant representations that are not intrinsic to mechanical systems [to do: mention some examples, e.g.
+in a footnote.
+The Lagrangian?
+Cartesian coordinates?].
+
+% Emphasize contrast: whereas North is arguing that we should believe that the state space of Hamiltonian mechanics corresponds to that of a classical world (using Occam's razor; inference to the best explanation), we are not trying to figure out which state space structure to believe in more.
+% As Swanson and Halvorson put it, North's argument presupposes ``that statespace structure actually corresponds to some kind of real, physical structure.
+Many physicists and philosophers of science view statespace as a formal tool for efficiently encoding certain modal commitments of scientific theories, and aid for making calculations and predictions, but not part of the theory's ontology.
+North's position presupposes some form of statespace realism" [1].
+North assuming that there is a fact of the matter regarding the state space of the world.
+
+Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as formal device for encoding physical information or even merely as a calculational tool.} Whether Hamiltonian phase space structure is metaphysically fundamental does not matter for the questions we seek to answer.
+Regardless of its metaphysical status, Hamiltonian mechanics is physically more fundamental, and we will argue for its epistemic and conceptual primacy as well.
+[To do: only keep preceding remark if somewhere later in the essay, we elucidate these epistemic and conceptual dimensions.] Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references].
+Each of these frameworks has different pragmatic costs and benefits.
+Hence, it is plausible that there may be no single, unique necessary mathematical structure for mechanics, but rather a family of related structures.
+There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions.
+Furthermore, each of these formalisms references some ``unphysical" coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.
+%The mathematical structure that you ``need" might not be unique.
+Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others.
+Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics.
+In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.
+
+North also introduces a technical notion of comparative structure, based on the symmetry group of statespace.
+According to this criterion, statespace A has more structure than statespace B if the symmetry group of A is a proper subgroup of B's symmetry group \parencites[87-88]{North}.\footnote{This statement follows the charitable reconstruction of North's criterion given by \textcites[]{Swanson}.
+North sometimes speaks in terms of comparing the dimensions of symmetry groups, which they note leads to trivial counterexamples.} A smaller symmetry group entails that it is in some sense ``harder" to preserve the space's structure, indicating that it has more mathematical structure to be preserved.
+These symmetry groups are defined in terms of the coordinate transformations on the space that preserve geometric structure, amounting to re-coordinatizations (redescriptions) of that structure.
+For Lagrangian systems, North takes the relevant symmetry group to be that of the point transformations, and for Hamiltonian systems, she assumes that canonical transformations play this role.
+Since point transformations are a proper subgroup of canonical transformations, North's criterion entails that the statespace of Lagrangian mechanics has more structure than the statespace of Hamiltonian mechanics---in virtue of the former having a smaller symmetry group.
+
+However, Swanson and Halvorson \parencites*[]{Swanson} develop a series of problems for North's argument.
+Most compellingly, they argue that we cannot in general assume that point transformations and canonical transformations are the proper symmetry groups for Lagrangian and Hamiltonian mechanics, respectively.
+As they note, Galilean boosts are often symmetry transformations of Lagrangian systems (in the sense of preserving the equations of motion), but they are not point transformations.
+Additionally, the two-body (Kepler) problem admits canonical transformations that map physically distinct solutions to each other (specifically, solutions whose orbits possess the same energy but different eccentricity).
+Hence, we cannot straightforwardly compare the structure of Lagrangian and Hamiltonian mechanics in terms of the symmetry groups of their statespaces.
+
+Advantageously, our criterion for comparing Hamiltonian and Lagrangian mechanics avoids the criticisms that Swanson and Halvorson develop against North.
+For the sense in which Hamiltonian mechanics has less structure than Lagrangian mechanics is simply one of logical strength: a system must satisfy strictly fewer physical conditions in order to be Hamiltonian.
+Whereas to be Lagrangian, the system must also satisfy the assumption of motion equivalence.
+The assumption of motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.
+
+Nevertheless, based on the equivalence between the tangent bundle and cotangent bundle formulations of Lagrangian and Hamiltonian mechanics, one might wonder whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics \parencites[1182]{Barrett2}.
+For, if there were, then this formulation of Lagrangian mechanics might be more general than the cotangent bundle formulation of Hamiltonian mechanics.
+And thus, there would be an analogous sense in which Lagrangian mechanics is more fundamental than Hamiltonian mechanics, undercutting our argument.
+Fortunately, it is implausible that there could be a suitable generalization of Lagrangian mechanics.
+In going from the cotangent bundle formulation of Hamiltonian mechanics to the symplectic manifold formulation, we lose some globally-defined structure.
+This is analogous to moving from the space of globally-Euclidean manifolds to the space of locally-Euclidean manifolds.
+Yet, it is not clear how Lagrangian mechanics could remain well-defined while throwing away global structure from the tangent bundle.
+For the action seems to be inherently a globally-defined quantity, as an integral defined over a trajectory.
+
+[To do: figure out best place to slip in the following background information, which is relevant for the preceding objection:] Barrett shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics are categorically equivalent provided that we define categories of models over the tangent and cotangent bundle formulations, respectively \parencites*[1181-82]{Barrett2}.
+However, he also shows that if we consider the more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are not categorically equivalent \parencites*[1182-83]{Barrett2}.
+This parallels the construction of Hamiltonian mechanics in Section~\ref{derivation}.
+
+% Plausibly, move the following directly after our claim that our notion of fundamentality/structure comparison avoids the objections of Swanson/Halvorson against North's:
+
 % Way to motivate that the motion equivalence assumption further constrains the Hamiltonian: the Hamiltonian already defines a relationship between conjugate momentum and velocity (through ($dq/dt = \partial H / \partial p$) ).
 Motion equivalence provides a further constraint on this relationship, and hence a further constraint on the Hamiltonian.
-\\
 
@@
 
-[Explain how our notion of comparative structure avoids the objections of Halvorson/Swanson against North's comparative notion of amount of structure.] \\
-
-We do not endorse a metaphysical notion of fundamentality, on which a formalism is more fundamental provided that it is more joint-carving.
-For instance, North provides the following definition of structure: ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" \parencites*[66]{North}.
-While Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{distribution}), we will also argue that classical mechanics deploys physically significant representations that are not intrinsic to mechanical systems.
-Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.
-Whether Hamiltonian phase space structure is metaphysically fundamental does not matter for the questions we seek to answer.
-Regardless of its metaphysical status, Hamiltonian mechanics is epistemically, conceptually, and physically more fundamental.
-
-% Part of the following paragraph seems redundant given the preceding. Think about how to combine.
-Methodologically, we agree with North that it is important to identify invariants and more generally ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}.
-Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North}.
-However, whereas North proceeds mainly from the mathematical structure to claims about physical significance, we seek to interpret the mathematics by beginning with physically significant assumptions.
-We seek to determine which physical assumptions are necessary and sufficient for supporting a corresponding mathematical structure.
-This does not entail that additional mathematical structure is necessarily physically insignificant or surplus structure, just that it may correspond to an additional physically significant assumption that might apply to the system under consideration.
-As we will see, this characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system.
-Hence, whereas North views the Lagrangian structure as surplus or redundant (``excess, superfluous structure"), we take the Lagrangian structure as unnecessary but still physically significant, corresponding to an additional physical assumption that applies to some classical systems.
-
-Furthermore, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references].
+%Finally, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references].
 Each of these frameworks has different pragmatic costs and benefits.
@@
 
-Claim to think about: GC was arguing on 10-12 that you could handle classical systems in the Hamiltonian framework with just a differential manifold and no symplectic structure.
+% Claim to think about: GC was arguing on 10-12 that you could handle classical systems in the Hamiltonian framework with just a differential manifold and no symplectic structure.
 This would contradict North's indispensability claim that Hamiltonian symplectic structure is necessary for treating classical systems.
@@
 
-% North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory.
-Structure comprises those quantities that remain intact when we transform one allowable coordinatization of the laws into another" [78]
-
-% North definition of structure: ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" [66].
-So presumably, North would view the invariant phase space density as part of this fundamental structure.
-
-% North's epistemic method for figuring out the structure: ``look at what is invariant under allowable transformations"; this tells us what the structure is [73].
-
-\section{Derivation}
+\section{Deriving Hamiltonian and Lagrangian mechanics}
 \label{derivation}
 
+% [To do: figure out how to flag to the reader that they can skip/skim some parts of this derivation section without loss of understanding for later sections].
+E.g., `Readers who are more interested in the significance of invariant phase space densities or the geometric action principle can feel at peace skipping ahead.' But, presumably, part of the stuff on invariant densities is relevant for Section~\ref{density}?
+Figure out how to finesse this
+
+This section justifies the physical and mathematical relationships in Figure~\ref{diagram}.
 Sections~\ref{infinitesimal}-\ref{deterministic} show how to derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution.
 Section~\ref{motion} shows how systems that satisfy the further condition of motion equivalence can be treated within the formalism of Lagrangian mechanics.
+We thus demonstrate in detail our methodology of deriving  relevant mathematical structures from underlying physical conditions.
+
+% Is it worth emphasizing/highlighting any particularly important parts of the derivation up front?
 
@@
 
-[Connect this with Curiel equation 6.2 for the Hamiltonian.]
+[To do: Connect this with Curiel equation 6.2 for the Hamiltonian.]
 
-\section{The physical significance of invariant distributions}
-\label{distribution}
+\section{The physical significance of invariant densities}
+\label{density}
 
-North argues that Hamiltonian mechanics is superior because it requires strictly less structure than Lagrangian mechanics.
-While we agree that Hamiltonian mechanics does require less structure---in a different sense than North's account---this is not the only grounds for privileging it.
-Instead, the main grounds for privileging Hamiltonian mechanics comes from understanding physics in terms of invariant distributions.
-This allows us to provide a better physical interpretation of the action principle, which holds even when a Lagrangian cannot be defined for the system of interest.
-% Do we agree with the following claim defended by North: ``as far as we can tell, we need only symplectic structure to do classical mechanics" [74].
+We have now substantiated the claim of Section~\ref{privileging} that Hamiltonian mechanics requires less structure than Lagrangian mechanics, in virtue of requiring fewer physical conditions.
+While this provides one reason for privileging Hamiltonian mechanics, it is not the only one.
+A further---possibly even more important---reason stems from the relationship between Hamiltonian mechanics and invariant distributions.
+As we argue in this section, viewing physics through the lens of invariant distributions has much to commend it.
+In general, it provides a better understanding of classical systems.
 
-Hamiltonian mechanics is ``fundamental" because it shows that densities are coordinate independent quantities and therefore physical quantities.
-Canonical pairs (position and conjugate momentum) allows a simpler expression/comparison of densities over space.
-That is why statistical mechanics works best on phase space (we have a uniform measure over q/p).
+Section~\ref{action} provides a further application of invariant distributions, providing a geometrical interpretation of the action principle.
+
+Section~\ref{fundamentality} relied on a notion of fundamentality in terms of generality.
+Here, we consider an epistemic sense of ``fundamental." In this sense, Hamiltonian mechanics is ``fundamental" because it shows that densities are coordinate-independent quantities.
+Assuming that coordinate-independence is sufficient for being physically significant, this shows that these densities are physically significant.
+More precisely, they are physical quantities or properties of the system.
+
+[To do: expand on the following point, and try to eliminate the notion of ``simplicity" and replace it with a more epistemic notion.
+Ability to make comparisons of density, independently of which coordinate system you are in or knowledge of being in a ``good" coordinate system.
+Josh: see 11-30-20 meeting notes where I think we discussed this.]: Compared with Lagrangian mechanics, canonical pairs (position and conjugate momentum) allow a simpler comparison of densities over space.
+
+[To do: expand on the following point: ] This is why statistical mechanics works best on phase space (we have a uniform measure over q/p).
+
+% Point of the following paragraph is to draw a contrast with North, who also seeks to identify physically significant quantities in terms of coordinate-invariants.
+We are pointing out that coordinate-invariance is not (on our view) a necessary condition for physical significance.
 
@@
 
-Moreover: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
+Further support for the preceding point: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
 (In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).
-% It follows from the way in which position quantities fix the units that $\dot{x} = v $.
+
+[Not sure if/where the following claim is relevant: ] This difference in the number of assumptions corresponds to a difference in the number of commitments.
+When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?).
+In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.
+
+[Not sure if/where the following claim is relevant: ] It follows from the way in which position quantities fix the units that $\dot{x} = v $.
 It is built into the notion of a dynamical system that half of the state variables are derivatives of the other half.
@@
 \section{Interpreting Lagrangian mechanics}
+\label{Lagrangian}
 
-The action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
-It also holds when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
-Also shows that Lagrangian is ``unphysical" and depends on gauge.
-
-Upshot: can better understand Lagrangian mechanics as a special case of Hamiltonian mechanics.
-Explain how the geometrical understanding of the stationary action in extended phase space should meet any criterion of ``naturalness" or ``natural construction." Lagrangian mechanics is actually less natural because contingent on having a bijection between velocity and conjugate momentum.
+By analyzing classical systems in terms of their invariant phase space density, we gain new insights into Lagrangian mechanics.
+Section~\ref{action} shows how the action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
+We show that the action principle holds even when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
+Section~\ref{gauge} discusses how the Lagrangian is highly gauge-redundant, and therefore less physical than the Hamiltonian.
+This provides a further argument for privileging Hamiltonian mechanics in the interpretation of classical systems.
+Finally, Section~\ref{Curiel} considers Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics, which he alleges encodes the necessary and sufficient structure for capturing the physically significant kinematical constraints of classical systems.
+This raises a number of objections to our own position, which we rebut.
+In short, we argue that we have given an eminently natural construction of Lagrangian mechanics, viewed as a special case of Hamiltonian mechanics.
+Moreover, our argument shows that Lagrangian mechanics is in some sense less natural than Hamiltonian mechanics because the former is contingent on having a bijection between velocity and conjugate momentum.
 
@@
 \subsection{Gauge redundancy in Lagrangian mechanics}
+\label{gauge}
 
-Claim to clarify: motion equivalence allows us to understand the gauge redundancy of the Lagrangian.
+% Last page of Halvorson/Swanson article seems to broach this issue of gauge redundancy, although they seem to think it is possibly a positive feature of Lagrangian mechanics (a source of additional representational flexibility) rather than a defect.
+
+In some sense (to be made more precise), gauge transformations are more prevalent in Lagrangian mechanics than Hamiltonian mechanics.
+This is so despite the fact that the conjugate momentum $p $ is itself a gauge quantity.
+
+[To do: clarify the following claim:] motion equivalence allows us to understand the gauge redundancy of the Lagrangian.
 However, it is not true that the Lagrangian has these redundancies in virtue of the system satisfying motion equivalence.
 
-In some sense (to be made more precise), gauge transformations are more fundamental in Lagrangian mechanics than Hamiltonian mechanics.
-The conjugate momentum $p $ functions as a dummy variable in virtue of being a gauge quantity.
-From motion equivalence, the Lagrangian has more invariant structure than the Hamiltonian [not sure what this really means].
+[To do: clarify the following :] From motion equivalence, the Lagrangian has more invariant structure than the Hamiltonian [not sure what this really means].
 The Lagrangian has the advantage of being invariant under coordinate transformations, whereas the Hamiltonian is not a coordinate invariant.
 Instead, the Hamiltonian varies as the time component of a 4-co-vector.
-Nevertheless, the Lagrangian itself is physically problematic because it is a gauge-dependent quantity.
-The p variables are not of fundamental significance because we can perform a gauge transformation on the p's such that the q and $\dot{q} $ variables remain the same.
-For systems where motion equivalence holds, we are really measuring the q and $\dot{q} $ variables, indicating that this is a genuine gauge transformation of the momentum variables.
+Nevertheless, the Lagrangian itself is physically problematic because it is a gauge-dependent quantity [elaborate/illustrate this].
+
+[Clarify the following, if relevant :] But there is a sense in which on the extended phase space, the Hamiltonian constraint is an invariant, yielding a coordinate-invariant way of handling systems with Hamiltonian mechanics.
+
+Gauge redundancy of conjugate momentum: The $p $ variables are not of fundamental significance because we can perform a gauge transformation on the $p $'s such that the $q $ and $\dot{q} $ variables remain the same.
+For systems where motion equivalence holds, we are really measuring the $q $ and $\dot{q} $ variables, indicating that this is a genuine gauge transformation of the momentum variables.
 Hence, privileging Hamiltonian mechanics does not entail, pace North, that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
 Instead, position and kinetic momentum remain physically privileged, with conjugate momentum relegated to a gauge quantity (despite its role in the Hamiltonian formalism).
-% But there is a sense in which on the extended phase space, the Hamiltonian constraint is an invariant, yielding a coordinate-invariant way of handling systems with Hamiltonian mechanics. (?) 
+
+[To do: connect the following point with the former :] Symplectic geometry indeed does not care about the difference between q's and p's.
+Classical mechanics does.
+Misindentification of the math (symplectic geometry) with the physics (classical Hamiltonian mechanics).
 %Since the q and $\dot{q} $ variables fix the physical state, this indicates that transformed conjugate momentum is a gauge quantity. 
-%North argues that ``momentum is a truly fundamental quantity, on a par with position" [77].
+%North argues that ``momentum is a truly fundamental quantity, on a par with position" [77]. I think that Wallace 2003 article might make a similar point. Look for this.
 
+\begin{comment}
 Issue of the mathematical structure that is necessary/sufficient for Lagrangian mechanics:
@@
 
-This difference in the number of assumptions corresponds to a difference in the number of commitments.
-When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?).
-In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.
-
 %Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics.
 Curiel frequently appeals to naturalness and a distinction between natural vs.
-ad hoc/artificial constructions [figure out what his definition of ``naturalness" is, if he provides one].
+ad hoc/artificial constructions [he defines `naturalness', specifically a `natural coordinate system', in terms of a particular kind of coordinate transformation.
+See his equation 4.3, page 22 of the PDF].
 
@@
 Moreover, Riemannian manifolds with large symmetry groups are very special, and a generic Riemannian manifold has no nontrivial symmetries."
+%Seems to me that this theoretical result supports North's analysis, and could offer a response to Halvorson/Swanson counterexamples that appeal to Kepler problem and Galilean boosts.
+
+\end{comment}
 
 \subsection{Response to Curiel's criticisms of Hamiltonian mechanics}
+\label{Curiel}
 
+Curiel \parencites*[]{Curiel} argues that Hamiltonian mechanics is less physically significant than Lagrangian mechanics.
+He attempts to show that the Lagrangian framework represents classical mechanical systems in a \textit{natural} matter whereas the Hamiltonian formulation involves an artificial construction \parencites*[270]{Curiel}.
+His argument hinges on the claim that Hamiltonian mechanics requires an ad hoc condition that is imposed by hand, whereas Lagrangian mechanics follows from physically plausible principles.
+[Set this up properly: Curiel's definition of abstract classical system; Curiel's privileging of certain kinematical constraints, his equations 4.3; Hamiltonian form of equation 6.2; see below].
 Curiel restricts Hamiltonian mechanics to systems whose Hamiltonian takes the form (equation 6.2) $H =\alpha^{m n} p_m p_n + U (q_j) $.
 Here, the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q $.
-Curiel's argument in a nutshell:  ``Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
-Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2)" (306).
+The following provides a helpful summary of Curiel's argument against privileging Hamiltonian mechanics:
 
-Curiel argues that Hamiltonian mechanics is less physically significant than Lagrangian mechanics.
-His argument hinges on the claim that Hamiltonian mechanics requires an ad hoc condition that is imposed by hand, whereas Lagrangian mechanics follows from physically plausible principles.
-We agree with Curiel that his equation 6.2 requires a condition that is imposed by hand, namely the kinematic equivalence assumption.
+\begin{quotation}
+Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
+Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2).
+\parencites[306]{Curiel}
+\end{quotation}
+
+We agree with Curiel that his equation 6.2 requires a condition that is imposed by hand, namely the motion equivalence assumption [explain this].
 However, this condition has a clear physical meaning, and it corresponds to a physical condition that must hold in order to define/apply Lagrangian mechanics.
@@
 
-Response: symplectic geometry indeed does not care about the difference between q's and p's.
-Classical mechanics does.
-Misindentification of the math (symplectic geometry) with the physics (Classical Hamiltonian mechanics).
-
-Curiel's main point: the representation of a classical mechanical system in the Lagrangian framework is \textit{natural} in a way that its representation in a Hamiltonian formulation is not.
+% Curiel's main point: the representation of a classical mechanical system in the Lagrangian framework is \textit{natural} in a way that its representation in a Hamiltonian formulation is not.
 The former representation depends on intrinsic structures of the classical mechanical system rather than an artificial construction (270).
 
-On its own, the motion of a system in physical space is not sufficient for defining the system's dynamics.
+Rebuttal 2: On its own, the motion of a system in physical space is not sufficient for defining the system's dynamics.
 The kinematics/motion does not suffice to determine the forces acting on the system.
@@
 
-The standard predator--prey model from differential equations satisfies Curiel's definition of an abstract classical system.
+Rebuttal 3: The standard predator--prey model from differential equations satisfies Curiel's definition of an abstract classical system.
 However, it does not satisfy Curiel's notion of the split between configurational vs.
@@
 
-Objection against Curiel to make precise and substantiate with textual evidence: implicitly, Curiel assumes that classical systems are defined as Newtonian systems in an inertial frame with Cartesian coordinates.
+Rebuttal 4: make precise and substantiate with textual evidence (might be relevant for responding to objection considered below): implicitly, Curiel assumes that classical systems are defined as Newtonian systems in an inertial frame with Cartesian coordinates.
 Implicitly, Curiel assumes that energy is a function of position and velocity, entailing that we have a Hamiltonian.
@@
 
+[Most sophisticated objection I can think of on behalf of Curiel:] Like our argument, part of Curiel's argument involves appealing to systems outside of the hyper-regular domain.
+Curiel acknowledges both that only hyper-regular Lagrangians have unique solutions and that Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}.
+Nevertheless, he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage.
+On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}.
+In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy a particular set of physically significant kinematical constraints (his equations 4.3).
+Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will still satisfy the kinematical constraints of Lagrangian mechanics \parencites*[311]{Curiel}.
+
+Summary: Hamiltonian mechanics provides solutions outside hyper-regular domain (but not these kinematical constraints) whereas Lagrangian mechanics preserves these kinematical constraints outside hyper-regular domain.
+Worry: perhaps GC and Curiel are in a verbal dispute about the proper/privileged definition of a classical system (with Curiel privileging these kinematical constraints, and GC privileging deterministic/reversible evolution).
+
+On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems).
+In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers the Hamiltonian $H=p$, which corresponds to a system moving with constant velocity (e.g.
+velocity of 1).
+So this could very well be a model of a photon, moving at the speed of light.
+Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating the Euler--Lagrange equation.
+Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310].
+But then does Curiel consider this system to not be a classical system?
+
+[To do: figure out if the following claims made by Curiel regarding the differences between the symplectic structure in Hamiltonian mechanics (which is fixed) vs.
+the symplectic structure that arises from the Euler--Lagrange equations (which varies from Lagrangian to Lagrangian) are relevant to our argument or rebuttals: \\
+Hamiltonian mechanics: ``in Hamiltonian mechanics, the fact that the symplectic structure is fixed once and for all enforces the satisfaction of the kinematical constraints (5.3), and it encodes the existence and uniqueness of solutions to the equations of motion and the conservation of energy, among many other things.
+All of these features are independent of the individual Hamiltonian used in the equation" [310].
+\\
+Lagrangian mechanics: ``In Lagrangian mechanics, in contrast, it is the Lagrangian itself that encodes existence and uniqueness of solutions (which will obtain when and only when the Lagrangian is regular), and there is not even a guarantee that energy is conserved, as one can, for example, represent dissipative systems in the Lagrangian framework" [310].
+
+% Either way, presumably we can agree with Curiel's explanation/understanding for why the Lagrangian must be regular: Curiel arguing that he can give an account of ``why it makes physical sense to demand that the Lagrangian be regular", i.e.
+satisfy equation 4.6 for the Hessian determinant.
+He notes that by considering the variational problem, we understand ``why the Lagrangian must be quadratic (at least) in the velocities for there to be a unique solution to the Euler--Lagrange equation" [308].
+Argument: for the variational problem to be well-posed, you need to be able to ``take non-trivial derivatives of the Lagrangian up to second-order in the velocity, since the crucial integration by parts that yields the Euler--Lagrange equation demands this, on pain of giving, for example, the tautology $0=0$" in some cases, or a contradiction in others.
+
 \section{Discussion of relation to categorical equivalence proofs}
 
-Barrett shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics are categorically equivalent provided that we define categories of models over the tangent and cotangent bundle formulations, respectively \parencites*[1181-82]{Barrett2}.
-However, he also shows that if we consider the more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are not categorically equivalent \parencites*[1182-83]{Barrett2}.
-This parallels the construction of Hamiltonian mechanics in Section~\ref{derivation}.
-
-% Not sure best place to put the following, if at all:
-Nevertheless, based on the equivalence between the tangent bundle and cotangent bundle formulations of Lagrangian and Hamiltonian mechanics, one might wonder whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics \parencites[1182]{Barrett2}.
-For, if there were, then this formulation of Lagrangian mechanics might be more general than the cotangent bundle formulation of Hamiltonian mechanics.
-And thus, there would be an analogous sense in which Lagrangian mechanics is more fundamental than Hamiltonian mechanics, undercutting our argument.
-Fortunately, it is implausible that there could be a suitable generalization of Lagrangian mechanics.
-In going from the cotangent bundle formulation of Hamiltonian mechanics to the symplectic manifold formulation, we lose some globally-defined structure.
-This is analogous to moving from the space of globally-Euclidean manifolds to the space of locally-Euclidean manifolds.
-Yet, it is not clear how Lagrangian mechanics could remain well-defined while throwing away global structure from the tangent bundle.
-For the action seems to be inherently a globally-defined quantity, as an integral defined over a trajectory.
-
-[To do: not sure if following warrants discussion:]
+[To do: not sure if following warrants discussion.
+Could possibly connect this with the earlier discussion of the significance of defining the unit system, since that seems key to our response to this hypothetical worry.]
 Teh and Tsementzis \parencites*[]{Teh} provide another argument for the equivalence of Lagrangian and Hamiltonian mechanics within the hyper-regular domain.

### 2020-12-14 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Minor comments and points to develop for sections 4 and 5.2

@@
 Understanding the relationship between Hamiltonian and Lagrangian mechanics depends partly on how one defines ``classical systems." Different definitions can result in different interpretive relationships.
-For instance, restricting classical systems to the hyper-regular domain precludes certain interpretive considerations, To avoid verbal disputes, it is essential to clarify what falls within the scope of ``classical systems." One way to delimit classical systems is to specify a set of equations that they satisfy---such as Hamilton's equations or the Euler--Lagrange equations.
+For instance, restricting classical systems to the hyper-regular domain precludes certain interpretive considerations.
+To avoid verbal disputes, it is essential to clarify what falls within the scope of ``classical systems." One way to delimit classical systems is to specify a set of equations that they satisfy---such as Hamilton's equations or the Euler--Lagrange equations.
 We will instead follow a different approach by delimiting classical systems based on the physical conditions that they satisfy.
@@
 We will view dissipative systems as Newtonian, but neither Lagrangian nor Hamiltonian.
+
+% General decomposition of conservative and non-conservative: https://scholar.harvard.edu/files/rsyuan/files/ijbc.pdf
 
@@
 Furthermore, since particle states are independent of a choice of units, this distribution must be invariant.
-We show in Section~\ref{infinitesimal} that any infinitesimally reducible system has a phase space given by a symplectic manifold.
+We show in Section~\ref{infinitesimal} that invariant distributions can only be defined over the phase space of classical mechanics, a symplectic manifold.
 The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
@@
 This is because knowing a system's position and velocity is not in general sufficient for knowing its momentum.
-For instance, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from only its trajectory through space.
+For instance, knowledge of the physical trajectory of a photon, when treated as a particle, is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from only its trajectory through space.
 In general, different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
@@
 Hence, the Jacobian determinant must in fact equal one.
-In other words, $\rho_c$ is a function only of the point $c$ and does not depend on the choice of coordinates.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\hat{\xi}^b)$ is the distribution defined over the transformed state variables, and $\rho_c(s(\hat{\xi}^b))$ is the distribution defined over the points of phase space, where $s: \mathbb{R}^n \to \mathcal{S}$ an invertible function  mapping a particular value of the units $\hat{\xi}^b$ to a point in the particle phase space $\mathcal{S}$.} We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
+In other words, $\rho_c$ is a function only of the particle state $s$ and does not depend on the choice of coordinates.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\xi^a)$ and $\rho_c(\hat{\xi}^b)$ represent the same distribution as expressed over the different coordinates and $s(\xi^a)$ and $s(\hat{\xi}^b)$ represent the same particle state as expressed by the different coordinates.
+Note that $\rho_c(s(\hat{\xi}^b))$ shows how the distribution depends on the coordinates only through the particle state.} We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
 In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a density and scalar function.
@@
 The evolution is therefore a symplectomorphism and corresponds to Hamiltonian evolution.
-Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes which leads to Hamiltonian evolution.
+Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes (i.e.
+sets of states are mapped to sets of states of equal size) which leads to Hamiltonian evolution.
 In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.
@@
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).
-Hence, weak equivalence suffices for the existence of a Lagrangian.
+Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.
 
@@
 \label{density}
+
+* "Fundamentality" because it makes the physics more prominent.
+
+* Invariant densities: physical significance connected to coordinate independence.
+
+* Density comparison is independent of coordinates. We can compare densities point-wise without worry.
+
+* Contrast this with densities over position and velocity.
+But!
+We can understand some aspects of Riemmanian geometry in terms of how states densities are represented over position and velocity.
+Hyper-regularity allows us to express densities from "states" to "kinematic variables".
+The fact that this expression depends only on position (and not on velocity), tells us we have a Riemannnian manifold.
+Norm of the metric tensor gives us the ``conversion'' factor.
+Note that norm of the metric tensor in general depends on the coordinate system and is in general point-dependent.
+There always exists a coordinate system in which the norm of the metric tensor is constant.
+The conversion factor for marginal distributions is linked to the eigenvalues along each coordinate.
+Flatness means there exists a coordinate system where we can factorize the distribution (if it factorizes in phase space), which is exactly the Cartesian coordinate system.
+(Note, we can always factorize in phase space) -> Therefore this focus on densities also allows us to get a better intuition even on the tangent space.
+
+* Makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
+Areas/volumes are measures of number of states.
+They are coordinate invariant.
+Also the Shannon/Gibbs entropy are coordinate independent (see physicality of quantities that are coordinate invariant)
+
+** Not everything physical has to be coordinate invariant.
+
+* Units are physically significant.
+In fact, the ability to express the densities in units is what gives us the symplectic structure.
+It's a physical requirement that we must be able to express the density in some unit.
+
+* Not all "state variable" are physical.
+Hamiltonian mechanics does not really care what the q's and p's are.
+It is the motion assumption that tells us that q's are position, simply because they define the unit system.
+P is a guage quantity (we care about the trajectory, the velocity).
+In this view, p is simply a ``label'' we can use to express the density over states.
+Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves.
+-> Invariant density is really the important physical element/object in Hamiltonian mechanics.
 
@@
 
+(XXX: Note how the geometry is better understood in extended phase space, requiring Hamiltonian picture.
+Note that this applies even when $\theta S$ cannot be expressed as function of $q$ and $\dot q$.
+)
+
 %Discussion of gauge transformations in the context of Lagrangian vs.
@@
 \label{gauge}
+
+** Overall point: this approach allows us to get a better intuition for what gauge transformations represent.
+
+* From motion equivalence -> we really only care about the trajectory. That is what we want to represent and what we measure.
+
+* A gauge transformation is a change in p variable that leave q and q dot unchanged.
+Therefore it is simply a change of how states are labeled, without a change in the actual observable dynamics.
+Add the derivation.
+
+* In Hamiltonian mechanics, the momentum is gauge dependent.
+In Lagrangian mechanics the kinematic variables are not gauge dependent, but the Lagrangian manifestly is.
+$L = \theta S$.
+$S$ is gauge independent but $\theta$ is not.
+Therefore we will have problems understanding the physicality of the Lagrangian.
+Similar concerns for the Action.
+-> neither formulation has the upper hand
+
+* Coordinate transformation: if we restrict to coordinate transformation that only depend on space, both Lagrangian and Hamiltonian are invariant.
+If we introduce coordinate transformation that do depend on time (e.g.
+rotating frame), then Lagrangian is still invariant while Hamiltonian manifestly changes as a time component of a co-vector.
+(because H is the time component of $\theta$).
+That is why the Lagrangian framework is prevalently used in intro classes.
+
+* This is a problem of the double role of time as ``coordinate'' and ``affine parameter''.
+To be proper in differential geometry, we have to separate the two uses.
+So we add a ``temporal'' degree of freedom $(t, E)$ and the evolution is over the affine parameter $s$.
+In that framework, the Hamiltonian constraint serves both as a global constraint that fixes the energy as a function of $q, p, t$ and as the generator of the affine parameter $s$.
+The Hamiltonian constraint is invariant.
+If the affine parameter is proper time, then the Hamiltonian constraint is essentially the mass.
+-> There is a better formulation of Hamiltonian mechanics, that is typically not found in textbooks, that addresses the problem.
 

### 2020-12-18 Articles/2019-Entropy/Source/thermodynamics.tex
Creating separate article for thermodynamics

\documentclass[letterpaper,twocolumn]{article}

\input{../../../include/basicstyle}
\input{../../../include/theorems}
\input{../../../include/logic}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\usepackage[font=scriptsize]{caption}

\begin{document}

\title{Counting evolutions: a simple foundation for thermodynamics}
\author{Gabriele Carcassi, Christine A. Aidala, et al \\ University of Michigan}

\date{\today}

\maketitle

\begin{abstract}
	In this work we present the core ideas at the heart of a new approach to the foundations of thermodynamics. The premise is to focus on the count of evolutions: the possible ways the state of a system can evolve in time. We see that in a deterministic process the evolutions can never diverge and in a reversible one they can never merge. This means that the count of evolutions per state can never decrease under a deterministic process and it is maximized at equilibrium. We show how this simple idea can be used as a straightforward foundation of thermodyanmics and how it relates to other notions of entropy in physics.

	This work is part of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
\end{abstract}

\section{Introduction}

This draft summarizes the status of our work on a new approach to the foundations of thermodynamics and is intended to gather early feedback. Rather then derive thermodynamics bottom-up from a particular microscopic model, we want to derive it top-down from a set of requirements on the dynamics of the system. The results we find are, then, independent on the microscopic dynamics in the sense that they can in principle be realized in different ways. This is both more in line with the aims of the founders of thermodynamics and justifies why similar concepts have been successful in fields beyond physics. Moreover, thermodynamics is found as a specialization of a more general model, which we believe can be used as a common foundation for the different types mechanics.

In a nutshell, we characterize processes in general terms, as the description of all possible ways that the system can evolve in time, its possible evolutions, under a particular set of circumstances, as one can see in figures \ref{fig_single_evolution} through \ref{fig_with_equilibria}. Counting the possible evolutions plays a fundamental role, in much the same way that counting states plays a role in statistical mechanics. In a deterministic and reversible process, all evolutions that start in one state will end in the same state: the evolutions cannot merge or diverge. If the process is deterministic but not reversible, however, the same final state can be reached from different initial states, and the evolution will merge. If the deterministic process reaches an equilibrium, there will be no more merging. We define process entropy to be the logarithm of the count of evolutions per state, which we find to have the basic properties we associate with entropy: it is linear under system composition of independent systems, can never decrease under a deterministic process, increases if the process is non-reversible and is maximized at equilibrium. This gives us a conceptually crisp physically meaningful concept that is tied to irreversibility and equilibria in a very natural way.

There are many technical details that are required to make the above characterization formally rigorous, mostly due to the proper handling of infinities. However, those do not really change the overall intuition, which we believe is the main strength of the approach. Therefore, in this paper, we will only give a conceptual overview leaving the formal details to other. This will leave the key physical insight is more directly accessible, and also result in a narrative that can be used directly in an introductory course in thermodynamics.

\section{States, processes and evolutions}

The first question we want to answer is: how can we characterize processes in the most general terms? We assume we have a system and we want to study its behavior in a given set of circumstances. If the state space $X$ of the system represents all possible configurations the system can be found in, ultimately we want to characterize how the state changes in time. As depicted in figure \ref{fig_single_evolution}, we have a copy of the state space at each moment in time, and we call an \textbf{evolution} a trajectory $x(t)$ that for each moment in time returns the state of the system.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{images/Slide1.png}
	\caption{The three sets represent the state space of the same system at three different moments in time. The blue line is an evolution, which tells us the state of the system at all times.}\label{fig_single_evolution}
\end{figure}

Not all evolutions will be possible under all conditions. Therefore we choose to characterize a \textbf{process} simply as the set of allowed evolutions, as we can see in figure \ref{fig_process}. Namely, a process tells us what can and cannot happen in a given set of circumstances. We also assume we have a measure $\mu$ that allows us to determine the size of a set of evolutions, to ``count'' evolutions.\footnote{The precise conceptual and mathematical definition for $\mu$ is source of a number of technical considerations which we (will) address in a separate work. While some issues, like unit definition, are important to the physics, most have to do with the proper handling of infinities or the conditions for existence and uniqueness of the measure. Overall, they would distract from the key insights.} A state $x_t$ at a time $t$ identifies a set of of compatible evolutions $A(x_t)$, those that pass through that state at that time. That is, $x(t) = x_t$. Using $\mu$ we can give a size to that set, which we can write as $\mu(A(x_t))$ or $\mu(x_t)$ for short.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{images/Slide2.png}
	\caption{A process is characterized by the set of all possible evolutions. The measure $\mu$ allows us to express the size of a set of evolutions. The probability of transitioning from one state to another corresponds to the fraction of evolutions that starts in the first and also ends in the second.}\label{fig_process}
\end{figure}

Note that the probability of transitioning from one state to another during a process is linked to the allowed evolutions and their count. In fact, calculating the probability of finding the state $x_{t+\Delta t}$ at $t+\Delta t$ given $x_t$ at $t$ means determining the fraction of evolutions that pass through $x_t$ that also pass through $x_{t+\Delta t}$. That is:
\begin{equation}
	P(x_{t+\Delta t} | x_t) = \frac{\mu(A(x_t) \cap A(x_{t+\Delta t}))}{\mu(A(x_t))}.
\end{equation}
Conceptually, the ability to count evolutions is equivalent to knowing all possible probability of transition.\footnote{Technically, a single probability measure is not enough to define all these transition at all scales, but the overall intuition remains.}

Having answered the question of how we characterize processes, we can categorize them. For example, we say a process is \textbf{deterministic} if the state at one time in enough to determine the state at all future times. We can see an example of such a process in figure \ref{fig_determinism}. In this case, if two different evolutions happen to reach the same state at a given time, they will also share all future states: evolutions can merge but cannot diverge. The count of evolutions per state over time can never decrease: $
\mu(x(t)) \leq \mu(x(t + \Delta t))$.

Deterministic processes can be characterized by a law of evolution. That is, we can write $x(t+\Delta t) = f(x(t), t)$ given a suitable function $f$ which may, in general, be time dependent. We can do this because $P(x(t+\Delta t) | x(t)) = 0$ for all but one future state. The law of evolution identifies that state. That is, $P(x(t+\Delta t) | x(t)) = 1$ if and only if $x(t+\Delta t) = f(x(t), t)$.

\begin{figure}[h!]
	\includegraphics[width=\columnwidth]{images/Slide3.png}
	\caption{Deterministic process: the state one time determines the state at future times. Evolutions can merge but not diverge in time. The number of evolutions per state cannot decrease.}\label{fig_determinism}
\end{figure}

Conversely, we say a process is \textbf{reversible} if the state at one time is enough to reconstruct the state at all past times. We can see an example of such a process in figure \ref{fig_reversibility}. In this case, if two different evolutions happen to reach the same state at a given time, they will also share all past states: evolutions can diverge but cannot merge. The count of evolutions per state over time can never increase: $
\mu(x(t)) \geq \mu(x(t + \Delta t))$.

\begin{figure}[h!]
	\includegraphics[width=\columnwidth]{images/Slide4.png}
	\caption{Reversible process: the state one time allows one to reconstruct the state at past times. Evolutions can diverge but not merge in time. The number of evolutions per state cannot increase.}\label{fig_reversibility}
\end{figure}

Note that in some literature the ability to reconstruct the past is called retrodictability and does not coincide with thermodynamic reversibility (the ability to undo a change). This distinction will be made precise and clear in the full framework. At this point we want to make it clear that our notion of reversibility will indeed correctly recover thermodynamic reversibility. It will do so without having to assume the existence of a reverse process or a symmetry under time-reversal.

A process that is both deterministic and reversible, like the one in figure \ref{fig_detrev}, will never contain evolutions that split or merge. The count of evolutions per state over time will be a constant: $\mu(x(t)) = \mu(x(t + \Delta t))$.

\begin{figure}[h!]
	\includegraphics[width=\columnwidth]{images/Slide5.png}
	\caption{Determinism and reversibility. Evolutions cannot merge nor split. The number of evolutions per state cannot increase.}\label{fig_detrev}
\end{figure}

To finish, we focus on deterministic processes with equilibria. We can see an example of such a process in figure \ref{fig_with_equilibria}. An equilibrium for a process is a state that, once it is reached, it is never left. In a process with equilibria all evolutions will reach, at some point, an equilibrium. In the out-of-equilibrium phase, evolutions merge until they reach equilibria and can no longer merge. This means that our measure $\mu$ is maximized at equilibria. If we care to characterize only the initial and final state, deterministic evolution simply means that the final equilibrium is a function of the initial state.

Note that the idea of equilibrium as ``nothing changing'' used here is an oversimplification, which does not even work relativistically. Yet, it suffices for now as it captures the intuition correctly.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{images/Slide6.png}
	\caption{Deterministic process with equilibria. Evolutions keep merging until equilibrium is reached. The number of evolutions per state increase and the maximum is reached at equilibria.}\label{fig_with_equilibria}
\end{figure}

Note that we have found a quantity, $\mu(x(t))$, that can never decrease in a deterministic process, stays the same if the process is reversible and is maximized at equilibrium. We have done so with no assumption on the type of system, the type of forces acting on the system, the internal composition of the system or the internal dynamics of the system. It knows nothing of uncertainty, disorder, information, statistical or probability distribution, phase space volumes. It only knows about evolutions and how to count them. The result is completely general and the justification is very straight forward. It is conceptually a perfect candidate for the thermodynamic entropy, expect for one thing we still need to understand: how it combines under system composition.

Suppose we now have two systems, with their respective state space $X$ and $Y$. The state space of the composite system is the set of all possible pairs $X \times Y$. A process defined on the composite system will come with a measure $\mu_{X \times Y}$ that enables us to count evolutions $\{x(t), y(t)\}$ of the composite system. In general, the state of one system may affect the other system, in which case the evolution of the composite cannot be separated. If the systems are independent, however, the evolution of one system will not influence the other. That is, for each possible evolution $x(t)$ of the first system and $y(t)$ of the second system, there will be a possible evolution $\{x(t), y(t)\}$ on the composite. In other words,
\begin{equation}
	\mu_{X \times Y} = \mu_X \mu_Y
\end{equation}
the measure factorizes. We define the \textbf{process entropy} as $S = \log(\mu)$. Since the logarithm is a monotonic function, the process entropy will preserve monotonicity and maximization of $\mu$ under deterministic process with equilibria. Which means:
\begin{equation}\label{ov_entropy_increases}
S(x(t)) \leq S(x(t + \Delta t))
\end{equation}
In addition, the process entropy will be additive for independent systems. As before there are a number of technical details that need to be addressed, most of all the units for $\mu$ which will influence the 0 point for $S$, which we will address in the full framework.

In general, the process entropy will depend on the process, not just on the state. However, a unique value for process entropy can be assigned to states of equilibrium. Suppose, in fact, that the state $x$ is an equilibrium for two different processes $A$ and $B$. Suppose that $A$ is at equilibrium $x$ and we switch to process $B$. Since $x$ is also an equilibrium for $B$, every evolution of $A$ that ended in $x$ will correspond to an evolution in $B$ that ends in $x$. By symmetry, the converse is also true. Therefore the process entropy of an equilibrium does not depend on the process, only on the equilibrium state. This allows us to assign a unique process entropy to the state. If we have set of states that are equilibria, then, we call \textbf{state entropy} $S(x)$ the process entropy that the system reaches at equilibrium.

To summarize, we found, under most general conditions, a quantity called process entropy, defined as the logarithm of the number of evolutions, that cannot decrease under deterministic processes, cannot increase under reverersible processes, is maximized at equilibrium and is additive for independent systems. Moreover, if $X$ consists of equilibria, there exists an \textbf{equation of state} $S : X \to \mathbb{R}$ which returns the state entropy, which is the process entropy of the equilibrium.

\section{Thermodynamics}

We now have all the basic tools we need to reconstruct thermodynamics, which we characterize as the study of equilibria of composite systems. This is not done in general, but under additional assumptions and definitions that tell us how the equilibrum of the whole relates to the equilibria of the parts. As such, the laws we find are not necessarily universal in nature: they apply to systems that satisfy (or can be assumed to approximately satisfy) the prerequisites.

We say a system is in \textbf{thermodynamic equilibrium} if:
\begin{itemize}
	\item it is at equilibrium
	\item all its parts are in thermodynamic equilibrium
	\item if another system joined with a part of the system in thermodynamic equilibrium is in thermodynamic equilibrium, then the whole is in thermodynamic equilibrium.
\end{itemize}
We say two or more systems are in thermodynamic equilibrium if their composite is in thermodynamic equilibrium. A \textbf{thermodynamic process} is a process with equilibria where the final state is a thermodynamic equilibrium. We again stress that we are not saying that two systems, in general, always allow for thermodynamic equilibria.

To get a feel for the definition, let us rederive the fact that thermodynamic equilibrium is transitive, also known as zeroth law. Suppose system $A$ is in thermodynamic equilibrium with system $B$ and system $B$ is in thermodynamic equilibrium with system $C$. Then the composite $B \times C$ is in thermodynamic equilibrium by definition. Since $A$ is in thermodynamic equilibrium with $B$, a part of $B \times C$, it is also in thermodynamic equilibrium with $B \times C$. Therefore $A \times B \times C$ is in thermodynamic equilibrium. But $A \times C$ is a part of $A \times B \times C$, so it is in thermodynamic equilibrium. Which means $A$ is in thermodynamic equilibrium with $C$.

The characterization of thermodynamic equilibrium we give is manifestly a part-whole relationship. It extends the notion of equilibrium to a composite system so that the equilibrium of the whole and the equilibrium of its parts are essentially the same equilibrium. We find that this kind of part-whole relationships are a good starting point not only for thermodynamics, but also for the other branches of physics. This should not come as a surprise as reductionism (i.e. explaining the whole through its parts) plays an important role in science.

We define a \textbf{thermodynamic system} one whose equilibria can be identified by a set of extensive quantities (i.e. additive under system composition). Moreover, one such quantity is called internal energy which has the property of remaining conserved under any process. We will denote $U$ as the internal energy and $x^i$ as the remaining extensive quantities. For a thermodynamic system, then, the state entropy can be expressed as $S(U, x^i)$ a function of the extensive quantities. This is simply the equation of state in entropic form, which is the starting point of Gibbs formulation of thermodynamics. [CITE] We can therefore define the relationships
\begin{align}
	\beta &= \frac{\partial S}{\partial U} = \frac{1}{k_B T} \;\;\;\;\;\; - \beta X_i = \frac{\partial S}{\partial x^i}
\end{align}
where $k_B$ is Boltzmann's constant. We find:
\begin{align}
dS &= \frac{\partial S}{\partial U} dU + \frac{\partial S}{\partial x^i} dx^i \\
&= \beta dU - \beta X_i dx^i \\
dU &= k_B T dS + X_i dx^i
\end{align}
which are the usual thermodynamic relationships, except that, for us, the state entropy $S$ dimensionless..

Note, again, that the system is identified by quantities that have a clear rule under system composition. We suspect that the existence of extensive quantities can be derived from the composability of thermodynamic equilibria. In fact, we typically construct measurement scales of quantities (such as energy, mass, volume, distance, ...) by preparing unit samples and then adding or dividing them. So it may be possible that, once we assume the ability to combine systems into bigger systems, we will consequently be able to construct extensive quantities without extra assumptions.\footnote{XXX: any work related to this?} Nevertheless, we leave these considerations for later works.

%Under these assumptions/definitions, we can recover the laws of thermodynamics. Let $A$ and $B$ be two thermodynamic system in thermodynamic equilibrium and let $AB$ be the composite. We have:
%\begin{equation}
%\begin{aligned}
%U_{AB} &= U_A + U_B \\
%S_{AB}(U_{AB}) &= S_A(U_A) + S_B(U_B) \\
%&= \sup \{S_A(x) + S_B(y) \, | \, x+y= U_{AB}\} \\
%\frac{1}{kT_A} = \beta_A = \frac{\partial S_A}{\partial U_A} &= \frac{\partial S_B}{\partial U_B} = \beta_B = \frac{1}{kT_B}
%\end{aligned}
%\end{equation}
%which follows the standard textbooks derivations and assumes the new equilibrium is not on a boundary point. In this case, equilibrium is achieved if and only if $T_A$ and $T_B$ are equal. Since equilibria are unique, this will also tell us that all equations of states have to be concave in $U$. If $B$ is in thermal equilibrium with another system $C$ we also have $T_B$ is equal to $T_C$ and therefore $A$ and $C$ are also in thermal equilibrium. This recovers the zeroth law.
%
%There is, however, a problem. We assumed the equilibrium is not a boundary point. It could be that system $A$ reaches its lowest energy state at a finite $T_A$ while $B$ can go to lower temperatures. This would mean that $B$

To derive the first and the second law we need to give a definition for work and heat. We do so by defining two special thermodynamic systems. We say a thermodynamic system $R$ is a \textbf{thermal reservoir} if its equation of state can be sufficiently approximated by $S = \beta U = \frac{U}{k_B T}$. Internal energy is the only state variable and the only way the entropy of the system can change. We call \textbf{heat} the energy exchanged by a thermal reservoir and we note $Q = - \Delta U$ as the energy lost by the reservoir during a thermodynamic process.

We say $M$ is a mechanical system if the state entropy is a constant for all states. For a thermodynamic system that is also mechanical, we therefore have  $S(U, x^i) = k_M$ for all its states. A mechanical system cannot change entropy and all states can be connected through a deterministic and reversible process. We have:
\begin{equation}
\begin{aligned}
dU &= k_B T dS + X_i dx^i = X_i dx^i.
\end{aligned}
\end{equation}
The energy of the mechanical system is effectively \emph{stored} in one of the other state variables and can be later retrieved. We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U$ as the energy gained by the system during a thermodynamic process.

These two types of systems allow us to study the two ``orthogonal'' directions a change of energy can have: an energy change in the direction of constant entropy and one in the direction of maximal energy change. These definitions makes this manifestly clear. Also note that $\frac{\partial S_M}{\partial U} = 0 = \beta_M = \frac{1}{k_B T_M}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite. While this ``temperature'' has little to do with thermometers, the fact that we can always transform all work into heat but not vice-versa is just a specific case of being able to move energy from higher to lower temperature.

We now consider a system composed of a generic system $A$, a thermal reservoir $R$ and a mechanical system $M$. During a thermodynamic process, because the total energy of the system must be conserved, we have:
\begin{equation}
\begin{aligned}
&\Delta U_A + \Delta U_R + \Delta U_M = 0 \\
&\Delta U_A - Q + W = 0 \\
&\Delta U_A = Q - W.
\end{aligned}
\end{equation}
This recovers the first law.

We now look at the change in entropy, which during a thermodynamic process can never decrease. Keeping in mind $M$ is mechanical, we have:
\begin{equation}
\begin{aligned}
&\Delta S_A + \Delta S_R + \Delta S_M \geq 0 \\
&\Delta S_A \geq - \Delta S_R \\
&\Delta S_A \geq \int \beta_R dQ \\
&\Delta S_A \geq \int \frac{dQ}{k_B T_R}.
\end{aligned}
\end{equation}
which recovers the second law of thermodynamics. This is just a more specialized version of \eqref{ov_entropy_increases} which is valid for any deterministic process.

The third law has many equivalent formulations. [CITE] One of them states that the entropy of any system has a non-negative lower bound which is reached by the lowest energy state. If the ground state is non-degenerate, such as a perfect crystal, its entropy is equal to zero. We can easily recover it in this form.

As state entropy counts the possible evolutions of the system at equilibrium, the lowest possible count $\mu$ is one, which is achieved if the equilibrium can be reached only by one starting condition. In this case, $S=\log(\mu)=0$. This tells us that state entropy is non-negative and therefore is bounded from below. We now have to show that all thermodynamic systems must have a lowest energy state and that it must be a local minima for entropy.

First of all, the existence of unique thermodynamic equilibria forces $S(U)$ to be strictly concave (i.e. negative second derivative). This means that, at least in some region, $S(U)$ must either be a decreasing or increasing function. But since $S$ cannot be negative, $S(U)$ cannot be defined on the whole domain. Therefore $U$ must be bounded at least from above or below. Consequently, the minimum for entropy is reached either at the lowest or highest energy state. Now we show that all thermodynamic systems must have either a lowest energy state or a highest energy state. Suppose, in fact, we had two system, $A$ has no maximum energy while $B$ has no minimum energy. Since $S_A(U_A)$ is concave, it must be a monotonically increasing function of $U_A$. This means that $\beta_A$ is always positive. Conversely, $S_B(U_B)$ must be monotonically decreasing, which means $\beta_B$ is always negative. If we put the two system in contact, the energy of $B$ will keep decreasing and the energy of $A$ will keep increasing and no equilibrium will ever be reached. Therefore all thermodynamic systems either have a minimum or a maximum energy. At this point the difference is a sign convention, so we can assume all thermodynamic systems have a minimum energy.

To finish, we note that a perfect crystalline structure fits the case where the system has only one possible evolution, so it is a zero entropy state. If there is degeneracy, there are different ways the equilibrium can be realized, there are different evolutions that will correspond to the same equilibrium which means the entropy will be greater than zero. This recovers the third law.

The laws of thermodynamics have been recovered, and we have done so only by characterizing how equilibria and state variables behave under system combination. Apart from that, we have made no assumptions on what type of system, the type of forces acting on the system, the internal composition of the system or the internal dynamics of the system. No reference to statistical mechanics or quantum mechanics is needed: the theory stands on its own. This fits the original spirit of the thermodynamics, which was to find a set of laws that would be independent on the ultimate dynamics of the constituents. Additionally, the concepts we used are directly related to the physical explanation, which is not recovered after pages and pages of calculation. The combination of these features makes this approach directly useful, for example, in an introductory physics course: a coherent self-contained explanation can be given in terms of simple definitions.

% \subsubsection{XXX Null states and thermodynamic zero}

% * Use null state as thermodynamic zero.
% * Must have zero entropy.
% * All extensive quantities for the null state must be zero.
% * Extensive quantities can be zero for a propert state (i.e. number of particle of a particular species).
% * Intensive quantities are derivative: they also have an absolute zero.
% * All extensive quantities must be non-negative?

% Sample text

%Zero entropy is usually associated to a non-degenerate ground state and the lower bound on energy is usually thought as a consequence of quantum mechanics. We saw how, in our framework, these requirements stem from the definitions. In fact, we can repeat the argument and find that \emph{all} extensive quantities. We also note that, while one typically focuses on temperature and entropy having an absolute scale, other thermodynamics variable, such as volume and pressure, have a well defined zero. Is there a better way to frame these issues? We believe there is.

%System composition $\times$ is a closed operation (i.e. given any two systems returns another system) that is commutative (i.e. $A \times B = B \times A$) and associative (i.e. $A \times (B \times C) = (A \times B) \times C$). Mathematically it makes sense to ask whether it has an identity (i.e. $A \times 0 = 0 \times A = A$), which we note as $0$ since the operation is commutative (i.e. additive). In fact, while the state space of the composite is the product of the subsystems, system composition in terms of systems is better thought of as an addition: we are putting things together. The identity, then, is naturally the ``null'' system: the empty system. We can keep adding nothing, and the system is unchanged. This makes system composition a commutative monoid, but not a group since there is no ``inverse'' system: the one that combined with the original gives us nothing.\footnote{Note that particles/anti-particle annihilation leaves something: the energy. So the ``anti''-system is not the ``inverse'' system.}

%The null system has only one possible configuration, which we call the null state, so it can only have one possible evolution. The null state of a null system is inherently in equilibrium and the state entropy is zero. In fact, the values for all extensive quantities must be zero since composing the null system with another null system still gives us a null system. Therefore, for example, $U_0 + U_0 = U_0$.

% The intensive quantities are defined as derivatives, which means their zero is also absolute: is the case where changing the extensive quantity leads to no change in entropy. This means that, in thermodynamic, all quantities must be absolute, not only temperature or entropy is absolute, but this tells us that all thermodynamic quantities must be absolute. Volume, pressure, number of particles, we already think of them as having a well defined zero.

%Some confusion comes from energy. In mechanics energy in the form of the Hamiltonian does not have an absolute zero, an arbitrary constant can be added with no ill effect. But rescaling internal energy by a constant $\hat{U}$ would make it no longer an extensive quantities: $(U_A + \hat{U}) + (U_B + \hat{U}) = U_{AB} + 2 \hat{U} \neq U_{AB} + \hat{U}$.\footnote{XXX: someone must have already said this} Now the question is whether

\section{Discussion}

We have shown that thermodynamics can be conceptually rederived by adding few other assumptions/definitions from a general characterization of states, processes and equilibria. Can this approach also offer new insights on classical mechanics, quantum mechanics and statistical mechanics that may lead to a more unified understanding these somewhat disconnected disciplines?

The notion of equilibria would be a place to start. We defined, as it is customary, the state space of a thermodynamic system as consisting only of states of equilibria. We also defined a mechanical system as one for which $S$ is the same for all states. But we said the state entropy is uniquely defined only for state of equilibria. This implicitly assumed that all states of a mechanical systems are equilibria. Is this always true? And in what sense is this true?

Let us first consider the classical phase space of a particle. Qualitatively, we can imagine saying that a particle at rest is in equilibrium. A particle in uniform motion, though, does not seem in equilibrium: the distance is constantly increasing. Yet, there exists a frame in which it is at rest. So, in this a sense, all free particles are at equilibrium. Going one step further, in the presence of gravity, a particle at rest in one frame will not be in uniform motion in another. So it would seem that a particle under gravitational or apparent forces is at equilibrium. What is going on?  The issue here is that, as we noted before, the notion of equilibrium as ``nothing is changing'', which is the notion coming from dynamical systems, is not exactly aligned with the notion of equilibria we need for thermodynamics.\footnote{XXX: cite papers that note the difficulty of definiting temperature relativistically} So we need to answer two questions: what would be a more suitable definition? And why would that definition be tied to the notion of state?

In another work, we more thoroughly investigate characterizing equilibria in terms of system decoupling, which we briefly present here. By system decoupling we mean a process that makes the system independent from other systems, its evolution does not constrain and is not constrained by the evolution of other systems. As we saw before, independence was already critical for defining state entropy. But why should it be connected to states? We typically think of a state as the description of one system and, implicitly, only that system. If the system is not independent, if there are correlations, the description of that system tells us something about the other as well. So if we want a state to give the description of one system alone, we implicitly describing it in the case where the system is independent. System independence is not just a requirement for the process entropy to factorize: it's an implicit requirement to properly talk about a system and its state. But why would independence be tied to a notion of equilibria? The idea is that when we group together quantities into a system, this is not done arbitrarily. For example, we do not group together the position of one ball with the velocity of another. We group quantities that ``go together'', that have to be studied together. More precisely, the ability to define a system impinges upon the ability to prepare it and study it independently from the rest. If we are not able to then what we have is not really a system, just a set of random quantities. So, at least in line of principle, we must have a way to decouple the system from other systems. We must have a process that makes the system independent. Clearly, if the system is already independent, the process should do nothing. If states are descriptions of an independent system, they must be equilibria of these decoupling processes. In this sense, we can say that all states, thermodynamic equilibria or not, are equilibria. Thermodynamic equilibria extends this notion of independence at different level of scale.

Still, one may find strange associating a state entropy to mechanical system. And why would that be constant on all states? We should note that the idea that all states have the same entropy is already embedded in statistical mechanics. For quantum mechanics, in fact, the Von Neumann entropy of every pure state is zero. In classical mechanics, a uniform measure is used to calculate the accessible states; additionally the Gibbs entropy is invariant under translations in phase space. It is also embedded directly in the corresponding mechanics. In both theories, given to states, we can always (at least mathematically) find a deterministic and reversible evolution (i.e. a Hamiltonian evolution or a unitary evolution) that connects the two. So the idea that mechanical systems are associated with a constant entropy, though seldom explicit, can be already deduced from the currently established theories.

Another natural question is: how does the process entropy relates to the fundamental postulate of statistical mechanics, that entropy is the logarithm of the count of states?
Suppose we give the description of a system at a given time (e.g. the energy of the system is within a particular bound). This will identify a set of possible evolutions and therefore a value for the process entropy. Also, this is the sum (or integral) of the process entropy associated with each state not ruled out by that description: every evolution will have to go through one of those states at that time. Now suppose the system is, at that time, independent from the other. Then the process entropy for each state is the state entropy, so the process entropy for that description is the sum (or integral) of the state entropy. Additionally, suppose the system is mechanical, then the process entropy for the description is, up to a constant, the count of states. Entropy as the count of microstates is easily recovered.

Note that this can happen in different scenarios. We could have an independent mechanical system under a deterministic and reversible process. Then counting evolution is the same as counting states, up to a constant, at any moment in time. In another scenario, the process is not deterministic and reversible but is isoentropic, meaning that the process entropy is constant in time. If we have an independent mechanical system, its evolution will necessarily be isoentropic since, whatever state we evolve to, the state entropy will always be the same. Now suppose the set of states compatible with our description mix with each other and only with each other. Each evolution will change state in time, but it will never go outside of the set, so the description works both as a constraint and a description of the dynamic equilibrium. At each time, the same description will correspond to the same states, so the process entropy will be constantly given by the count of states.

Then one may ask: how are these two entropies different? Imagine that the mechanical system starts independent but the dynamics is non-reversible leading the dynamical system to an equilibrium, as in a damped harmonic oscillator. The process entropy at the initial state will coincide with the count of states, but not at later times. As the evolution concentrate around the equilibrium, the number of states become fewer so the number of evolutions per state increases. The entropy of the equilibrium, assuming that it is ever reached, is essentially the size of the basin of attraction at the initial time. The entropy around the equilibrium, then, increases in time which is consistent with the process being irreversible. Conversely, note that the areas in phase space shrink as the available states become fewer and fewer, which would mean entropy as the log of the state count decreases under an irreversible process. It was actually this paradox that pushed into considering evolutions as more fundamental than states and, retrospectively, it is the obvious choice.

This example also show how critical is the notion of independence to define state entropy. The evolution of a damped harmonic oscillator is not independent from the evolution of the systems that are absorbing the dissipated energy. The same description at different times (e.g. the system has reached equilibrium) does not provide exactly the same information on the system or other systems. In these cases, the state entropy does not allow us to compare apples to apples while the process entropy always does.

Having seen how the process entropy relates to the count of states, one may naturally ask: how does it compare with the Gibbs/Shannon entropy? Suppose you cannot distinguish the dynamics of a system under a given time resolution. That is, in principle the system has a well defined trajectory $x(t)$ but you can only measure quantities of the form $A[x(t)] = \int_{t_0}^{t_0 + \Delta t} a(x) dt$ (e.g. average energy, average number of particles, ...). A permutation of states within the trajectory will not affect any of the measurement outcomes. Therefore the best we can do, given enough of those measurements, is to measure the frequency $\rho(x)$ each state is visited within the time resolution. Now we ask, given a specific distribution $\rho(x)$, how many possible evolutions $x(t)$ would that correspond to? That means counting the number of permutations $W[\rho]$ we can construct. The process entropy will be $\log W$. The Shannon/Gibbs entropy is the limit of that expression if we take $x(t)$ to be continuous.

The approach, then, is not simply a nicer retelling of thermodynamics, but it can help us better understand the fundamental connection between all these theories. We stress that we are not arguing that all other notions of entropy should be abandoned in favor of process entropy: quite the contrary. Specific tools are often more useful than general tools precisely because they take advantage of the specific nature of the problem. The point is that if we are more aware of the interplay between the general definition and the more specific one we will not use the latter inappropriately.

\section{Conclusion}

\bibliography{bibliography}

\end{document}

### 2020-12-21 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Wrote section 5.2 (gauge redundancy) and parts of Section 4 (significance of invariant density); implementing section outlines from last week.

@@
 \end{abstract}
+
+\tableofcontents
 
@@
 Furthermore, since particle states are independent of a choice of units, this distribution must be invariant.
-We show in Section~\ref{infinitesimal} that invariant distributions can only be defined over the phase space of classical mechanics, a symplectic manifold.
+We show in Section~\ref{infinitesimal} that to define this invariant distribution, we require a symplectic manifold, which serves as the phase space of classical mechanics.
 The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
@@
 
-* "Fundamentality" because it makes the physics more prominent.
-
-* Invariant densities: physical significance connected to coordinate independence.
-
-* Density comparison is independent of coordinates. We can compare densities point-wise without worry.
+* ``Fundamentality" because it makes the physics more prominent.
+
+* Invariant densities: physical significance connected to coordinate independence. [incorporated]
+
+* Density comparison is independent of coordinates. We can compare densities point-wise without worry. [incorporated]
 
@@
 (Note, we can always factorize in phase space) -> Therefore this focus on densities also allows us to get a better intuition even on the tangent space.
+[Still need to incorporate much of this discussion]
 
@@
 
-* Not all "state variable" are physical.
+*[Could save the following for discussion of gauge redundancy in section 5.2] Not all ``state variables" are physical.
 Hamiltonian mechanics does not really care what the q's and p's are.
@@
 
+[To do: is there a physically or conceptually important distinction between coordinate-invariance vs. coordinate-independence? (vs. co-variance?)]
+
 We have now substantiated the claim of Section~\ref{privileging} that Hamiltonian mechanics requires less structure than Lagrangian mechanics, in virtue of requiring fewer physical conditions.
 While this provides one reason for privileging Hamiltonian mechanics, it is not the only one.
-A further---possibly even more important---reason stems from the relationship between Hamiltonian mechanics and invariant distributions.
-As we argue in this section, viewing physics through the lens of invariant distributions has much to commend it.
+A further---possibly even more important---reason stems from the relationship between Hamiltonian mechanics and invariant densities.
+As we argue in this section, viewing physics through the lens of invariant densities has much to commend it.
 In general, it provides a better understanding of classical systems.
-
-Section~\ref{action} provides a further application of invariant distributions, providing a geometrical interpretation of the action principle.
-
-Section~\ref{fundamentality} relied on a notion of fundamentality in terms of generality.
-Here, we consider an epistemic sense of ``fundamental." In this sense, Hamiltonian mechanics is ``fundamental" because it shows that densities are coordinate-independent quantities.
+Specifically, we will see that invariant densities are of fundamental physical significance due to their coordinate-independence [invariance?].
+Hamiltonian mechanics---unlike Lagrangian mechanics---allows us to compare these densities point-wise in any coordinate system.
+Although the density is physically significant because it is coordinate-invariant, we will nevertheless argue that coordinate-invariance is not a necessary condition for physical significance.
+In fact, the ability to express densities in some choice of units is closely connected with symplectic structure.
+Hence, we will argue that some coordinate-dependent representational choices are nevertheless physically significant.
+This view is at odds with metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms [get citations: Field 1980; Chen papers; North].
+% Section~\ref{action} uses the invariant density to provide a geometrical interpretation of the action principle.
+
+In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
+If the scalar value of the density at one point is larger than another, this entails that the density is higher at that point of phase space.
+Physically, this means that there is more physical stuff at this point.
+Likewise, if the scalar value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
+For brevity, will refer to this as the property of \textit{point-wise comparison}.
+Since the density depends in general on the choice of units, point-wise comparison does not hold universally across representations of state space.
+Instead, for the density of states to have this property, we require a coordinate-independent definition of the density.
+As the derivation in Section~\ref{deterministic} shows, we must define our state space using a symplectic manifold in order to define an invariant density.
+In other words, to define a density with the property of point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
+As we will discuss further in Section~\ref{gauge}, the conjugate momentum $p$ is itself a gauge-dependent quantity.
+Hence, the real physical significance of canonical coordinates is that they allow us to define an invariant phase space density $\rho $.
+
+It is instructive to contrast this situation with Lagrangian mechanics.
+In this case, the density depends on the position and velocity state variables $(q,\dot{q}) $.
+Crucially, these state variables transform contravariantly, rather than covariantly.
+Consequently, the density $\rho(q,\dot{q})$ also varies contravariantly.
+These transformation properties preclude the density---in the Lagrangian formalism---from being coordinate-invariant in general.
+Of course, there exist special coordinate systems that allow for faithful point-wise comparison of the density.
+For instance, Cartesian coordinates---for a system in an inertial frame of reference---support this formal representational property.
+Yet, we still need to know which coordinate system we are in (or have transformed to) before making these point-wise comparisons.
+In general, non-Cartesian or non-inertial coordinate systems will not allow for point-wise comparisons of the density.
+The Hamiltonian formalism circumvents this further epistemic requirement by defining the density using quantities that vary covariantly.
+Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
+
+[Should we briefly illustrate or gesture at an illustration of a failure of point-wise comparison?
+E.g., consider a situation where the density is a constant $k$ as a function of Cartesian coordinates (x, y).
+If we transform to polar coordinates $(r,\theta) $, then the scalar value of the density is no longer a constant (due to the Jacobian of this transformation not being unity).
+For instance, the scalar value of the density increases as we go further out, even though the density is physically the same.
+Hence, to determine whether there is more stuff at one location in state space vs.
+another, we need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
+we can't just compare the values of $\rho$ in these coordinates and see which is larger).]
+
+% Connect this back with fundamentality: epistemic notion of the fundamental (don't need to know which coordinate system we are in), along with a physical sense: supporting a physical invariant, namely the invariant phase space density.
+Showing that this is more physically significant than the state variables themselves.
+
+Section~\ref{fundamentality} analyzed fundamentality in terms of generality.
+We argued that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in part because it is more general.
+By considering how Hamiltonian mechanics supports faithful point-wise density comparisons, we have seen two further notions of fundamentality.
+The first is epistemic: by representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
+Any and all coordinate systems will support the formal property of point-wise comparisons [what about time-dependent coordinate transformations?].
+Not only is this formal property methodologically advantageous, but it also bespeaks a further physical significance.
+This brings us to a second, physical notion of fundamentality: the phase space density $\rho$ is physically fundamental because it is a coordinate-invariant.
+Assuming that coordinate-invariance is sufficient for physical significance, the density thereby represents a genuine physical property of classical systems.
+It quantifies the amount of stuff in each phase space state.
+[to do: should we make this precise in terms of representational theory of measurement?
+E.g., the particular kind of scale that this representation defines and the meaningful comparisons it supports?].
+
+% Here, we consider an epistemic sense of ``fundamental." In this sense, Hamiltonian mechanics is ``fundamental" because it shows that densities are coordinate-independent quantities.
 Assuming that coordinate-independence is sufficient for being physically significant, this shows that these densities are physically significant.
@@
 
-[To do: expand on the following point, and try to eliminate the notion of ``simplicity" and replace it with a more epistemic notion.
+[To do: expand on the following point: ] This is why statistical mechanics works best on phase space (we have a uniform measure over q/p). See the relevant * point above.
+
+% Point of the following paragraph is to draw a contrast with North, who also seeks to identify physically significant quantities in terms of coordinate-invariants.
+We are pointing out that coordinate-invariance is not (on our view) a necessary condition for physical significance.
+
+While it is important to provide a physical interpretation of invariants, this is not to say that only invariant quantities are physically meaningful.
+For instance, the coordinates $q$ define the units, and units are physically significant despite not being invariant.
+Units are physically significant because they allow us to assign numbers to a physical quantity.
+They are physically significant in part because they correspond to measurable quantities.
+Indeed, as the derivation in Section~\ref{deterministic} shows, a necessary condition for defining the symplectic structure is that we can express the density in a particular choice of units.
+If we did not have the physical ability to define the density in some units---although any particular choice is arbitrary---we would not be able to define this structure.
+[To do: make this argument more clear]
+
+Whereas North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}, some choices of the coordinates $q$ are less physically meaningful because they are not properly connected with measurement.
+Some arbitrary mixings of $q$'s and $p$'s preserve the symplectic manifold but are not as physically significant.
+For this reason, the $q$'s are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
+Whereas symplectic geometry creates a mathematical, syntactic symmetry between the $q$ and $p$ variables, this is not a physically significant symmetry from the standpoint of measurement.
+%cite van Fraassen on the measurement point; e.g. APA session on Halvorson book
+
+Further support for the preceding point: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
+(In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).
+
+% Have mostly done the following: [To do: expand on the following point, and try to eliminate the notion of ``simplicity" and replace it with a more epistemic notion.
 Ability to make comparisons of density, independently of which coordinate system you are in or knowledge of being in a ``good" coordinate system.
 Josh: see 11-30-20 meeting notes where I think we discussed this.]: Compared with Lagrangian mechanics, canonical pairs (position and conjugate momentum) allow a simpler comparison of densities over space.
-
-[To do: expand on the following point: ] This is why statistical mechanics works best on phase space (we have a uniform measure over q/p).
-
-% Point of the following paragraph is to draw a contrast with North, who also seeks to identify physically significant quantities in terms of coordinate-invariants.
-We are pointing out that coordinate-invariance is not (on our view) a necessary condition for physical significance.
-
-While it is important to provide a physical interpretation of invariants, this is not to say that only invariant quantities are physically meaningful.
-For instance, the coordinates q define the units, and units are physically significant despite not being invariant.
-Units are physically significant because they allow us to assign numbers to a physical quantity.
-They are physically significant in part because they correspond to measurable quantities.
-Whereas North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}, some choices of the coordinates q are less physically meaningful because they are not properly connected with measurement.
-Some arbitrary mixings of q's and p's preserve the symplectic manifold but are not as physically significant.
-For this reason, the q's are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative p's are not.
-Whereas symplectic geometry creates a mathematical, syntactic symmetry between the q and p variables, this is not a physically significant symmetry from the standpoint of measurement.
-
-Further support for the preceding point: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
-(In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).
 
@@
 
-[Not sure if/where the following claim is relevant: ] It follows from the way in which position quantities fix the units that $\dot{x} = v $.
-It is built into the notion of a dynamical system that half of the state variables are derivatives of the other half.
+% By definition, a dynamical system is one where we can write xdot as a function of x (i.e.
+the current state determines the evolution of the system).
+Having a Hamiltonian system further means that we can express this function xdot(x) using the Hamiltonian.
 
@@
 
+\begin{comment}
+
+% 12-19: I think that I incorporated the relevant points below:
+
 ** Overall point: this approach allows us to get a better intuition for what gauge transformations represent.
@@
 
+\end{comment}
+
 % Last page of Halvorson/Swanson article seems to broach this issue of gauge redundancy, although they seem to think it is possibly a positive feature of Lagrangian mechanics (a source of additional representational flexibility) rather than a defect.
 
-In some sense (to be made more precise), gauge transformations are more prevalent in Lagrangian mechanics than Hamiltonian mechanics.
+Understanding the action principle in terms of Hamiltonian mechanics also provides a better understanding of gauge transformations.
+If motion equivalence holds, then gauge transformations leave the variables $(q,\dot{q}) $ invariant.
+To see this, recall that motion equivalence entails that knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
+In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$.
+Hence, any transformation that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant, i.e.
+is a gauge transformation.
+[To do: make it more clear how the discussion in this section is intrinsically connected with the earlier geometric treatment of the action and our overall privileging of Hamiltonian mechanics]
+
+In particular, there is a family of gauge transformations that changes the conjugate momentum variable $p$ while leaving $(q,\dot{q}) $ invariant.
+[To do: illustrate this with general formula for this kind of gauge transformation, and possibly the electrodynamics example.] Consequently, the conjugate momentum in Hamiltonian mechanics is gauge-dependent.
+The $p $ variables are therefore not of fundamental physical significance [to do: check that this is consistent with Section 2.1's discussion of fundamentality].
+Hence, privileging Hamiltonian mechanics does not entail, \textit{pace} North, that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
+[To do: see if Wallace 2003 also makes this claim about the fundamentality of conjugate momentum, or the similarly flawed claim that conjugate momentum and position are on equal footing within Hamiltonian mechanics].
+Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.
+
+Instead, position and kinetic momentum are physically privileged, since these quantities are gauge-invariant.
+Consequently, Lagrangian mechanics employs state space variables that are gauge-invariant, while Hamiltonian mechanics uses a gauge-dependent variable.
+Despite this advantage, Lagrangian mechanics has a compensating disadvantage: the Lagrangian is itself gauge-dependent.
+The formalism of Section~\ref{action} makes this gauge dependence manifest.
+Representing the Lagrangian as $L =\vector{\theta} \cdot \vector{S} $, we see that since the vector potential $\vector{\theta}$ is gauge-dependent while the displacement vector $\vector{S}$ is not, $L$ must also be gauge-dependent.\footnote{As shown in Section~\ref{action} we can arbitrarily transform $\vector{\theta}$ by adding the gradient of an arbitrary scalar function, while $\vector{S}=curl(-\vector{\theta}) $ remains invariant under this gauge transformation.} The gauge-dependence of the Lagrangian makes it difficult to provide a clear physical interpretation of any particular Lagrangian.\footnote{The action inherits the gauge-dependence of the Lagrangian.
+This does not affect the principle of stationary action because the local minima and maxima of the action remain invariant under gauge transformations.} In contrast, the Hamiltonian is defined up to a constant, endowing it with greater physical significance.
+Overall then, we see that neither Lagrangian nor Hamiltonian mechanics has the upper hand when it comes to invoking only gauge-independent physical quantities.
+Both formalisms have their trade-offs in this regard (although, it is worth emphasizing that---as argued in Section~\ref{density}---Hamiltonian mechanics provides the natural setting for the invariant phase space density).
+
+Moving from gauge transformations to coordinate transformations, we gain further reasons for using Hamiltonian mechanics within the extended phase space formalism.
+For many applications, it suffices to use time-independent coordinate transformations.
+In this special case, both the Lagrangian $L$ and the Hamiltonian $H$ are coordinate-invariant.
+However, some applications benefit from (or even require) the use of time-dependent coordinate transformations, e.g.
+a rotating frame of reference.
+In this case, the Lagrangian remains coordinate-invariant, while the Hamiltonian becomes coordinate-dependent.
+Again, the formalism of Section~\ref{action} makes this time-dependence manifest.
+As we saw there, the Hamiltonian $H$ is the time component of the vector potential $\vector{\theta} $, entailing that $H$ transforms as the time component of a co-vector.
+Methodologically, this makes the Hamiltonian framework less convenient for treating this class of problems.
+
+Nonetheless, formulating Hamiltonian mechanics within the extended phase space provides one way to rectify these deficiencies.
+At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and an affine parameter within the usual Hamiltonian framework.
+Yet, differential geometry demands that we separate these two roles.
+We thereby add a temporal degree of freedom $(t, E) $, expanding the phase space degrees of freedom by one (recall that we previously had one degree of freedom for every canonical coordinate pair $q_i, p_i$).
+Next, we define the phase space evolution over an affine parameter $s $.
+In this extended Hamiltonian framework, the Hamiltonian constraint functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [to do: spell out what the `Hamiltonian constraint' is in general].
+Crucially, the Hamiltonian constraint is now coordinate-invariant, even with respect to time-dependent coordinate transformations.
+For example, if we define the affine parameter as the proper time, then the Hamiltonian constraint is proportional to the mass (which is manifestly coordinate-invariant).
+This formalism thereby provides a clear physical meaning for the Hamiltonian constraint.
+[To do: provide some reference for this modified Hamiltonian formulation]
+
+[To do: not sure how best to end this subsection/transition to the next one].
+We have now considered many conceptual and physical advantages of the Hamiltonian formalism.
+In Section~\ref{Curiel}, we turn to considering Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.
+
+[Could address the following in this section, rather than in Section 4:] Not all ``state variables" are physical.
+Hamiltonian mechanics does not really care what the q's and p's are.
+It is the motion assumption that tells us that q's are position, simply because they define the unit system.
+P is a gauge quantity (we care about the trajectory, the velocity).
+In this view, p is simply a ``label'' we can use to express the density over states.
+Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves.
+This entails that the invariant density is really the important physical element/object in Hamiltonian mechanics.
+
+\begin{comment}
+
+I think that I have incorporated most of the following into the above:
+
+I didn't incorporate the following point: In some sense (to be made more precise), gauge transformations are more prevalent in Lagrangian mechanics than Hamiltonian mechanics.
 This is so despite the fact that the conjugate momentum $p $ is itself a gauge quantity.
 
-[To do: clarify the following claim:] motion equivalence allows us to understand the gauge redundancy of the Lagrangian.
+% [To do: clarify the following claim:] motion equivalence allows us to understand the gauge redundancy of the Lagrangian.
 However, it is not true that the Lagrangian has these redundancies in virtue of the system satisfying motion equivalence.
@@
 [Clarify the following, if relevant :] But there is a sense in which on the extended phase space, the Hamiltonian constraint is an invariant, yielding a coordinate-invariant way of handling systems with Hamiltonian mechanics.
-
-Gauge redundancy of conjugate momentum: The $p $ variables are not of fundamental significance because we can perform a gauge transformation on the $p $'s such that the $q $ and $\dot{q} $ variables remain the same.
-For systems where motion equivalence holds, we are really measuring the $q $ and $\dot{q} $ variables, indicating that this is a genuine gauge transformation of the momentum variables.
-Hence, privileging Hamiltonian mechanics does not entail, pace North, that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
-Instead, position and kinetic momentum remain physically privileged, with conjugate momentum relegated to a gauge quantity (despite its role in the Hamiltonian formalism).
 
@@
 %North argues that ``momentum is a truly fundamental quantity, on a par with position" [77]. I think that Wallace 2003 article might make a similar point. Look for this.
+
+\end{comment}
 
@@
 North assumes that formulating a Lagrangian theory requires specifying a Riemannian metric; according to Curiel, this is neither necessary nor sufficient.
-Furthermore, the physical significance of a Riemannian metric in the context of a Lagrangian theory is unclear: two different Lagrangian's can have the same Lagrangian vector field and yet induce inequivalent Riemannian metrics
+Furthermore, the physical significance of a Riemannian metric in the context of a Lagrangian theory is unclear: two different Lagrangians can have the same Lagrangian vector field and yet induce inequivalent Riemannian metrics
 
@@
 
-\section{Discussion of relation to categorical equivalence proofs}
+\section{Material looking for a possible home}
 

### 2020-12-21 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Notes from discussion

@@
 * ``Fundamentality" because it makes the physics more prominent.
-
-* Invariant densities: physical significance connected to coordinate independence. [incorporated]
-
-* Density comparison is independent of coordinates. We can compare densities point-wise without worry. [incorporated]
+Some people use coordinate-free/independence as necessary conditions for fundemantaly/physical significance.
+This means coordinates do not appear in the definition.
+Yet, densities require coordinate for the definition, therefore cannot be coordinate-free/independent, therefore there are some physically significant quantities that do not satisfy that.
+Why fundamental???
+(reverse mathematics?
+fundamental because it recovers the correct math from the physical assumption?)
+
+---
+Section on densities and riemmanian geometry. (Gabriele)
+
+* A paragraph to show how densities work.
+In general are coordinate dependent and you can't simply compare values at different points.
+This is the case in Lagrangian mechanics because both differentials in position vary controvariantly.
 
@@
 
-* Makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
-Areas/volumes are measures of number of states.
-They are coordinate invariant.
-Also the Shannon/Gibbs entropy are coordinate independent (see physicality of quantities that are coordinate invariant)
-
-** Not everything physical has to be coordinate invariant.
-
-* Units are physically significant.
-In fact, the ability to express the densities in units is what gives us the symplectic structure.
-It's a physical requirement that we must be able to express the density in some unit.
+---
+
+Section on Ham. Not everything physical has to be coordinate independent or invariant.
+
+* In phase space, we do have coordinate invariance and point-wise comparison.
+Momentum changes covariantly.
+It is the density that is truly fundamental, not the coordinates in general.
+Though, the q's are priviledged from epistemic access.
+The p's are not, in fact we can define gauge transformation as things that keep the q's the same but change the p.
 
@@
 
-[To do: is there a physically or conceptually important distinction between coordinate-invariance vs. coordinate-independence? (vs. co-variance?)]
+* Units are physically significant.
+In fact, the ability to express the densities in units is what gives us the symplectic structure.
+It's a physical requirement that we must be able to express the density in some unit.
+
+* Invariant densities: physical significance connected to coordinate independence. [incorporated]
+
+* Density comparison is independent of coordinates. We can compare densities point-wise without worry. [incorporated]
+
+* Makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
+Areas/volumes are measures of number of states.
+They are coordinate invariant.
+Also the Shannon/Gibbs entropy are coordinate independent (see physicality of quantities that are coordinate invariant)
+
+* Close the section with a more general discussion on physicality/fundamentality and all those other physical goodness.
 
@@
 
+Understanding both mechanics in terms of our assumptions gives us a way to understand gauge transformations that is both more physically grounded and intuitive.
+Recall that motion equivalence entails that knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
+In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$.
+Hence, any transformation that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant.
+We want such transformation to be canonical as to preserve the invariant densities.
+This means that the new state variable $(q, \hat{p})$ will respect the usual Poisson brackets:
+\begin{equation}
+\begin{aligned}
+	\{q^i, q^j\} &= 0 \\
+	\{q^i, \hat{p}_j\} &= \delta^i_j \\
+	\{\hat{p}_i, \hat{p}_j\} &= 0
+\end{aligned}
+\end{equation}
+Using the second equation we find:
+\begin{equation}
+\begin{aligned}
+	\{q^i, \hat{p}_j\} &= \frac{\partial q^i}{\partial q^k} \frac{\partial \hat{p}_j}{\partial p_k} - \frac{\partial q^i}{\partial p_k} \frac{\partial \hat{p}_j}{\partial q^k} \\
+	&= \delta^i_k \frac{\partial \hat{p}_j}{\partial p_k} - 0 \frac{\partial \hat{p}_j}{\partial q^k} = \frac{\partial \hat{p}_j}{\partial p_i} = \delta^i_j
+\end{aligned}
+\end{equation}
+Integrating we have:
+\begin{equation}
+\begin{aligned}
+\hat{p}_j = p_j + A_j(q^i)
+\end{aligned}
+\end{equation}
+where $A_j$ are arbitrary functions of position. Using the third Poisson bracket, we have:
+\begin{equation}
+\begin{aligned}
+0 = \{\hat{p}_i, \hat{p}_j\} &= \{p_i + A_i, p_j + A_j\} \\
+&= \{p_i, p_j \} + \{p_i , A_j\} + \{A_i, p_j \} + \{A_i, A_j\} \\
+&= 0 - \{A_j, p_i\} + \{A_i, p_j \} + 0 \\
+&= - \frac{\partial A_j}{\partial q^i} + \frac{\partial A_i}{\partial q^j} \\
+&= curl(A_j) \\
+\end{aligned}
+\end{equation}
+Since $A_j$ is curl free, we can express it as the gradient of a function $f$.
+Therefore changes of state variable of the form $\hat{p}_j = p_j + \partial_j f$ do not change either the equations of motion or the trajectories.
+We recognize these as the gauge transformations.
+
 Understanding the action principle in terms of Hamiltonian mechanics also provides a better understanding of gauge transformations.

### 2020-12-23 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Adding section on density/riemmanian geometry

@@
 [Still need to incorporate much of this discussion]
+
+\subsection{Density and Riemmanian geometry}
+
+It is helpful to take a quick look at how densities work in general so that we can better understand what is so special about them in the context in Hamiltonian mechanics.
+By density we mean the intuitive physical notion of a quantity divided by region it is found, like mass per volume or charge per area.
+Over a continuum, this is an integrable function $\rho(x)$.
+Mathematically, the concept is captured by the Radon–Nikodym derivative: we have a region $U$, we have two measures $m$ and $V$ and the derivative satistifies $m(U) = \int_U \rho dV$.
+This makes it clear that the units of $\rho$ depends on the units of both $m$ and $V$.
+More precisely, if $\rho(x, y)$ is a function of multiple variables, we will want $\int_U \rho(x, y) dx dy$ so that the units of $\rho$ are consistent with the units of the arguments.
+For example, in polar coordinates $\rho(r, \theta)$ will return mass over meter-radians.
+
+This object is coordinate dependent because its definition requires us to make a choice of coordinates.
+Worse than that, it is not coordinate invariant: the density changes with the Jacobian.
+For example, changing polar to Cartesian we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
+Since the Jacobian is a function of the point, this makes the value of the density, in general, not directly comparable at different points.
+If we have a uniform density in $x$ and $y$, we will have an $r$ dependence in $r$ and $\theta$.
+This dependence is introduced precisely because the area identified by $dr$ and $d\theta$ is $dr \,r d\theta$, which is a function of $r$.
+
+We can look at this through the lens of differential geometry.
+First of all, note that in differential geometry there is also an object called density, but this is not the function $\rho$ we are looking for.
+A density in differential geometry is essentially a volume form (i.e.
+an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation.
+What we are looking for is really the component of a volume form $\nu = \rho(x^i) dx^1 \wedge dx^2 \wedge ...
+\wedge dx^n$.
+Note that $\nu$ combines the density and the volume, and that is why it is coordinate-free.
+Comparing the integrals we have $m(U) = \int \nu$ so essentially $\nu = dm$ which is the infinitesimal mass contribution we are integrating.
+But this means the value returned by $\nu$ does not depend only on how must stuff is there at the point, but on the size of the volume identified by the $n$ vectors given to the form.
+
+In Riemannian geometry, we have a canonical volume form $dV = \epsilon \sqrt{|g_{ab}|} dx^1 \wedge dx^2 \wedge ...
+\wedge dx^n$ where $|g_{ab}|$ is the determinant of the metric tensor as expressed in local coordinates and $\epsilon$ is $1$ or $-1$ depending on the orientation.
+It can also be expressed as $\sqrt{|g_{ab}|}\epsilon_{a_1 ...
+a_n}$ in component notation where $\epsilon_{a_1 ...
+a_n}$ is the Levi-Civita symbol.
+We can then express any differential density $\nu = \hat{\rho} dV$ as the product of the volume density times a scalar function $\hat{\rho}$.
+By simply looking at the math, we may be mislead to think we were able to express the density in a coordinate independent way, but this is not the case.
+The quantity $|g_{ab}|$ depends on the choice of coordinates and is, in general, a function of the point.
+However, we can always find a set of coordinates that, at least locally, makes that quantity constant.
+The determinant $|g_{ab}|$ is the Jacobian of the transformation to those coordinates.
+For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$; in polar coordinates $|g_{ab}| = r^2$ and the volume form $dV = r dr \wedge d\theta$.
+This means that $\hat{\rho}$ is really expressing the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise.
+We can also see that if we look at the units of $\hat{\rho}$, which is terms of inverse volume as expressed in the privileged system.\footnote{This is similar to what happens with the concept of proper time: it is defined in the comoving frame, so it is coordinate dependent.
+Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate invariant.}
+
+We can now look at what happens for distributions over position and velocity.
+Note that the velocities transform in the same way (i.e.
+controvariantly) than differentials in position.
+Therefore $\rho(x^i, v^j) = \sqrt{|g_{ab}|} \sqrt{|g_{ab}|} \rho(\hat{x}^i, \hat{v}^j) = |g_{ab}| \rho(\hat{x}^i, \hat{v}^j)$ $dx^a$.
+The determinant of the metric tensor is therefore directly the volume of the tangent bundle.
+The relationship to the density as expressed in phase space is therefore the following:
+\begin{equation}
+	\rho(q^i, p_j) = |g_{ab}| \rho(x^i, v^j).
+\end{equation}
+A change of coordinates will induce a controvariant change in the velocity and a covariant change in the momentum.
+The left side will remain invariant.
+The two terms on the right size will change in the opposite way so that their product is invariant.
+Focusing on densities, then, brings about a physically motivated way to understand the relationship between the cotangent and the tangent bundle, where the metric tensor plays a role in defining the geometry of the second in a similar way that the symplectic form defines the geometry of the first: it defines the count of states so that densities can be properly tracked.
 

### 2020-12-23 Articles/2019-ShannonEntropy/Submissions/EJP/EJP-ShannonEntropyIsVariability.tex
Shannon article EJP submission

--- 2019-12-28 Articles/2019-ShannonEntropy/Submissions/AJP/AJP-ShannonEntropyIsVariability.tex
@@
-\documentclass[prb,preprint]{revtex4-1} 
+\documentclass[iopart]{revtex4-1} 
 % The line above defines the type of LaTeX document.
-% Note that AJP uses the same style as Phys. Rev. B (prb).
+% Note that EJP is an IOP journal and should use iopart.cls, but that iopart.cls is supposedly incompatible with amsmath.sty.
 
@@
 Defining a set of questions means choosing an encoding as we are choosing how the information gets codified into our series of bits.
-To understand how this works, we can briefly review the Huffman coding,[CITE] which is the optimal algorithm for symbol-by-symbol coding with a known probability distribution.
+To understand how this works, we can briefly review the Huffman coding,\cite{Huffman} which is the optimal algorithm for symbol-by-symbol coding with a known probability distribution.
 The idea is that we want all possible answers to each question to be balanced, to provide the same amount of information.

### 2020-12-29 Articles/2019-ShannonEntropy/Submissions/EJP/response.tex

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epsfig,amssymb,amsmath,amsthm}
%\usepackage[active]{srcltx}
%\usepackage[hypertex,linkcolor=red]{hyperref}

\usepackage{color}
\usepackage{cancel}
\usepackage{tikz-cd}

\DeclareMathOperator{\spn}{span}

\newcommand{\pj}[1] {\underbar{$#1$}}
%\newcommand{\pj}[1] {\overline{#1}}

\def\>{\rangle}
\def\<{\langle}
\def\ca{_{\cal A}}
\def\cb{_{\cal B}}
\def\cc{_{\cal C}}
\def\comment#1{}
%\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
\def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
\def\labell#1{\label{#1}}
%\def\labell#1{\label{#1}{\mbox{{\tiny #1}}}}
%\def\section#1{{\par\em #1:--- }}
\def\togli#1{}
\def\sh{\mbox{sh}}
\def\iden{\openone}

\begin{document}

	Thanks editor.

\section*{Referee 1}

\begin{quote}
Overall this paper is clearly written. This paper gives a comprehensive review of the Shannon entropy. The Shannon entropy characterizes the variability of the discrete random variables. This is widely known to an expert but this paper state this idea clearly with examples. This kind of heuristics and intuitive ideas are very valuable for newcomers.
\end{quote}
We thank the referee for the kind words and for appreciating the goal of the paper.

\begin{quote}
However, the authors miss the key reference: Mathematical foundations of information theory by Khinchin. The requirements on page 3 were rigorously presented in this book. 
\end{quote}
We added the reference.

\begin{quote}
The entropy is invariant under the transformation if and only if the transformation is volume-preserving. This statement should be clearly stated on page 6.
\end{quote}
We agree: we only highlighted unitarity of the Jacobian, which not everyone will intuitively connect to volume preservation. We also lacked a comment on marginal distributions. We revised the paragraph which now reads:

\emph{[...] This has a couple of important consequences. First of all, (1) the volumes are preserved during the transformation, which means (2) the Jacobian is unitary and (3) the Shannon variability is invariant. In fact, these three properties are mathematically the same property: each one implies all others. Second, something similar happens to the areas for each independent degree of freedom, meaning that the Shannon variability associated to the marginal distributions is also invariant. These properties can be taken to be the defining characteristics of phase space: phase space has exactly such geometrical properties.}

\begin{quote}
Another minor thing, on page 14, should be 13/12 split instead of 13/14.  
\end{quote}
We sincerely thank the referee for catching it: these types of minor mistakes are a source of distraction to newcomers.

\section*{Referee 2}

\begin{quote}
	The key conceptual flaw is that the authors clearly consider that Information is not a physical quantity;
\end{quote}

No such claim is made in the paper. The following claim is made:

\emph{The use of information in physics, then, does not warrant a fundamental change of perspective in what constitutes a physical object, as some physicists have claimed.}

\noindent which is not at all the same. This simply means, as it is explained, that while matter and energy carry information through their physical processes, they are not themselves ``made'' of information.

\begin{quote}
	even though they wrote in their introduction “we do not discuss the relationship [of the Shannon entropy] to the thermodynamic entropy, which is a separate concept,” in an apparently neutral and unpartisan view of this important issue.
\end{quote}
Thermodynamic entropy is defined in terms of temperature and heat. The paper does not talk about temperature and heat, which are essential if one wants to link thermodynamic entropy to Shannon entropy. That has nothing to do with information being physical.

TODO add footnote

\begin{quote}
However, since the underlying implicit thesis to their paper is that the Shannon information is not physical, they should justify this fundamental premise still further.
\end{quote}
That is not our thesis. We simply state, which is not a matter of personal opinion, that the nature of the Shannon entropy depends on the nature of the ``probability'' one is using. In ecology, for example, $p_i$ often represent the fraction of the population of animals that belong to the same species $i$. The Shannon entropy, which is called Shannon diversity index, is an objective (and physically grounded) quantity. In machine learning and Bayesian probability, the $p_i$ represent the degree of belief of an agent that something will happen. The Shannon entropy will represent, by definition, something subjective. We stated this very clearly in the original submission:

\emph{This expression quantifies the variability of the elements within the distribution, the variety of values one finds. The precise meaning of this variability is context dependent, as the choices of the elements and binning are not fixed and the meaning of the weights depends on what the distribution is describing. But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope, a probability and so on.}

TODO: add here that similar comments are throughout the paper

\begin{quote}
 Their main reference supporting this premise appears to be their ref.[4] – however, this particular reference [4] is itself highly flawed (having both numerous technical and conceptual errors, and also completely ignoring the highly relevant work of E.T. Jaynes – it also a paper not published in an appropriate physics journal – however, I’m not writing here a review of this flawed ref.[4] paper.) But, the fact that the authors of this present paper considered the ref.[4] to be an appropriate reference to support their fundamental assumption about the ‘unphysicality’ of information therefore also unfortunately casts a negative light on their scientific judgement.
\end{quote}
That reference appears after the sentence ``it is still surrounded by a great deal of confusion''. Therefore it is only used as evidence to show that there is confusion surrounding entropy. This can be taken as the reference showing the confusion, or the reference itself being confused, as the referee argues. In either case, it proves the point: there is confusion.

We find the attacks on the reference completely out of place, inappropriate and distasteful. Ad hominem attacks have no place in the scientific discourse.

\begin{quote}
Hence, their wish to use the word ‘variability’ as an alternative synonym to the word ‘entropy’.
\end{quote}

The referee, once again, puts words in our mouth that are not on the page. The use of `variability' for the Shannon formula is justified by this paragraph:

\emph{The first problem is that the formula is attached to different concepts that are specific to each field. [...] In this paper we show that the idea of variability of the elements within a distribution, how diverse the objects are from each other within the collection, can serve as such a general concept. We will therefore use the term \emph{``Shannon variability,''} leaving ``entropy'' to the physical concept.}

So, ``variability'' is used for something that works not only in physics, but also ecology, information theory, machine learning, communication theory, economics... ``Entropy'' is used for the physical concepts in statistical mechanics. That helps the reader understand that some properties of that concept are more general than physics. Given that we make a distinction between the two words, ``variability'' and ``entropy'' are \textbf{not} synonyms in our work. That is the whole point. The variability of species in ecology has nothing to do with entropy in statistical mechanics.

\begin{quote}
Rather, it is clear to me the Shannon formula is actually the fundamental physical quantity from which all entropy types (be it the information, Gibbs, and Boltzmann entropies etc.) can be derived.
\end{quote}
This claim is hotly contested in the literature of the foundations of thermodynamics. [cite] In modern textbooks of statistical mechanics, the Shannon entropy is not the starting point, rather an approximation that is valid only at equilibrium [cite]. The referee is therefore expressing his personal view.

We are staying out of that debate and we would find inappropriate, in the context of a paper aimed at scientific education, to take a side as the referee suggests.

TODO: add in the introduction

\begin{quote}
The current authors appear to have completely neglected the more recent and highly relevant work written by Jaynes in the 1970s, 1980s, and 1990s, but relied only on his original seminal papers of 1957 and 1965.
\end{quote}

This is beyond the scope of the paper, which is simply to introduce in a clear way to newcomers what the expression $- \sum p \, \log p$ represents.

\begin{quote}
In particular, the authors here complain that the information entropy is ‘subjective’ in contrast to thermodynamic entropy which is ‘objective’. However, their treatment of this is completely superficial. They need to address Jaynes’ more nuanced and deeper treatment of this issue in his 1992 paper.
\end{quote}
Again, we are in no way disproving what he says and the referee is engaging in what philosophers define as ``verbal dispute'', where one disagrees on the meaning of the words rather than on an actual fact of the matter. We mean `subjective' in the Bayesian/machine learning sense, where each agent has a different take on how much something is likely. Jaynes, unfortunately, used language that sometimes seems Bayesian to refer to cases where the experimental setup matters (which is not Bayesian). This language is then taken out of context and used to support claims that do not follow from Jaynes' examples. So our discussion is there to caution the target reader, who is a newcomer to all of this, of these differences.

Given that the referee keeps mis-characterizing our prose, it is difficult for us to understand what, if anything, gave him the wrong impression. Note that we tested the article on a group of students and none of these misunderstandings emerged.

TODO: review the 1992 paper, see if there is something about physicality

\begin{quote}
	There is an egregious technical error in this paper: In the discussion of the Continuous Distributions (Section IV) after Eq.1 the authors discuss the quantity log(b-a) and suggest that it can be negative for $0<b-a<1$. However, on the assumption that b and a (i.e. the co-ordinate x) in general represent some kind of physical quantity, e.g. length [m]) you simply cannot take the logarithm of a dimensioned quantity! A normalising unit quantity, e.g. $\Delta$, which also defines the granularity or resolution is required to normalise the argument of the logarithm; i.e. we need log\{ (b-a)/$\Delta$ \}. The choice of the granularity (which is a key aspect to any entropic analysis) automatically solves this apparent problem raised by the authors. But they have overlooked even this simple (mathematical and physical) aspect.
\end{quote}
Let $\Delta > b - a > 0$. Then $\frac{b-a}{\Delta} < 1$. Therefore $\log \frac{b-a}{\Delta} < 0$. The resolution does not fix the possible negative value of the Shannon entropy for continuous distributions. The importance of granularity, of unit choice, not only was already present in the introduction, but is literally given in the following paragraph of the original submission:

\emph{	When considering continuous distributions, we have to remember that the densities are given per unit. [...] A distribution over half a unit interval will correspond to minus one bit, since we would need to ``lose'' one bit of information to widen the variability back to one unit. In other words, the variability is measured by the number of bits needed to identify an element up to one unit.}

Again, given the complete mis-characterization of the prose and the erroneous claim that somehow the choice of unit/granularity makes the Shannon entropy always positive, we find it impossible to understand whether there is any meaningful item to address.

TODO: add "granularity"

\begin{quote}
Actually, the authors clearly show their general lack of understanding of the importance of granularity in entropy considerations (especially entropy essentially being a scale-less quantity) since they consider the Shannon entropy to be somehow ‘perplexing’ since it changes when the scale is stretched, shrunk, or changed. However, this is simply equivalent to a change of granularity (or scale) where it is well known that the calculated entropy will therefore always concomitantly change. 
\end{quote}
Again, it is difficult to understand whether, hidden under all the ad hominem, there is an actual point. The word `perplexing' is used in the following paragraph:

\emph{	This should be particularly perplexing if we want to use this quantity in a physical setting. If we want $\rho(x)$ to represent the state of a macroscopic system and $H(\rho(x))$ to represent entropy, a state variable, how can $H$ change value if we merely express $\rho$ over different coordinates? How can maximization of entropy be meaningful if it yields different results depending on the coordinate system? Does that mean that we can use the Shannon formula only on distributions over discrete values?}

The word `perplexing' is used as a rhetorical device, to introduce questions a newcomer might find puzzling, which we answer in the following paragraph. Does the referee really think that means we find the issue perplexing? This would leave us very perplexed...

\begin{quote}
The authors comment that “infinitesimally narrow distributions, like delta functions, do not work well with the Shannon [entropy]”. It is well known that delta functions always require extreme care in their handling, since they are intrinsically unphysical (not conforming to many of the key requirements of physically plausible functions, such as causality and the absence of infinities etc.); so to suggest that the Shannon entropy is somehow inappropriate due to its having apparent problems with the delta function is fatuous – it just means that the delta function is not being appropriately applied. 
\end{quote}
Again, if the referee actually read what we wrote, he would have seen that we say exactly the same thing: that one has to work with functions that have small but still finite support.

\begin{quote}
The authors make a curious comment about the Boltzmann constant, that it “is really just set by the unit system”, as if it isn’t a universal constant and is simply a matter of convention! 
\end{quote}
In the last revision of the SI, universal constants \textbf{are} set to selected values, no longer measured: they are a matter of convention. They are called \textbf{defining constants} because they define the unit system, which is a convention. Before engaging in comments that imply we are incompetent, the referee may want to read the relevant literature.

TODO: add footnote

\begin{quote}
There are numerous other contentious and erroneous (careless) statements made by the authors. I’ve merely focussed on the larger, fundamental issues to indicate that this paper is not fit for publication.

Overall, the authors have created a ‘strawman’ argument (by using only partial and weak sources for their argumentation) against the Shannon entropy, so that their paper makes no useful contribution into the literature for a more fundamental understanding of entropy.
\end{quote}

--------------------

Closing.

Sincerely,

The Authors

\end{document}

### 2020-12-29 Articles/2019-ShannonEntropy/Submissions/EJP/EJP-ShannonEntropyIsVariability.tex

@@
 
-The Shannon entropy $-\sum p_i \log p_i$ is well established in such disparate fields as computer science, physics\cite{Jaynes1957-1, Grandy} and ecology, yet it is still surrounded by a great deal of confusion.\cite{Swendsen, KishFerry} For example, in statistical mechanics it is often introduced and associated with vague notions (such as uncertainty, knowledge, lack of knowledge and disorder) which prevent a proper understanding of the concept.\cite{Styer, Ben-Naim} In this work we concentrate on giving a better characterization of the Shannon entropy and its relation to statistical mechanics by addressing what we believe are the two main problems; we do not discuss the relationship to the thermodynamic entropy, which is separate concept.
+The goal of this work is to give a comprehensive introduction to the concept of Shannon entropy, the expression $-\sum p_i \log p_i$, which is well established in such disparate fields as computer science, communication theory, physics\cite{Jaynes1957-1, Grandy} ecology and economics.
+The general concept of entropy in physics, despite being treated extensively in the physics education literature (e.g.
+[cite physed]), is still surrounded by a great deal of confusion [cite evidence of confusion]\cite{Swendsen, KishFerry}.Partly because, at a fundamental level, there is still disagreement on what is the ``correct'' notion of entropy, particularly out of equilibrium.
+[cite works of out of equilibrium research]  We will focus on the notion of Shannon entropy, which has received less attention in physics education.
+We will see that the Shannon entropy can be given a crisp and precise characterization and will explore its relation to the Gibbs and Boltzmann entropies in statistical mechanics.\footnote{We will not discuss its relation to thermodynamic entropy: this requires the notions of heat and temperature, which will not be covered in this work.}
 
-The first problem is that the formula is attached to different concepts that are specific to each field.
-To disentangle this issue we need to provide a characterization of the Shannon entropy that satisfies the following requirements:
+To give a crisp and intuitive meaning to the Shannon entropy we need to address two main problems.
+The first problem is a consequence of its success.
+As it is used in many fields, it is generally introduced with concepts that are specific to that field but may not be appropriate for others.
+As we will see, the meaning of the $p_i$ can vary significantly.
+The first goal, then, is to provide a characterization of the Shannon entropy that satisfies the following requirements:
 \begin{itemize}
@@
 In this paper we show that the idea of variability of the elements within a distribution, how diverse the objects are from each other within the collection, can serve as such a general concept.
-We will therefore use the term \emph{``Shannon variability,''} leaving ``entropy'' to the physical concept.
+We will therefore use the term \emph{``Shannon variability''} for this more general idea that applies to all cases, leaving ``entropy'' to the more specific physical concept.
 We will show that the expression $- \sum_i p_i \log p_i$ is the only linear indicator of variability and it measures the number of questions one must ask to identify an element of the distribution, linking its use to information theory.
 The variability of messages within a communication channel will correspond to its information entropy as used in communication theory; the variability of microstates within a macrostate will correspond to the Gibbs entropy used in statistical mechanics; the variability of the state of a single particle within a microstate will correspond to the Boltzmann entropy.
+This avoids the use of incomplete or imprecise characterizations (e.g.
+uncertainty, knowledge, lack of knowledge, disorder) that, while useful in some limited circumstances, promote confusion in others.
 
 The second problem is that the Shannon formula presents some peculiarity when applied over continuous variables that is often the source of confusion.
-Namely, it is generally linked to a choice of unit.
-Yet, the physical state spaces, in both classical and quantum mechanics, provide additional structure that make the Shannon variability coordinate-independent.
+Namely, it is generally linked to a choice of unit, of the resolution used to describe the problem.
+However, the physical state spaces, in both classical and quantum mechanics, provide additional structure that make the Shannon variability coordinate-independent.
 This is seldom pointed out, and it is critical to form a proper connection to the Boltzmann, Gibbs and von Neumann entropies.
@@
 These are the same requirements Shannon put forth for his expression\cite{Shannon}, from which he showed that the only possible choice is $H(p_i) = - \sum p_i \log(p_i)$.
+See also \cite{khinchin2013mathematical} for a similar derivation.
 This expression quantifies the variability of the elements within the distribution, the variety of values one finds.
@@
 If we have binary questions, the logarithms will be in base two and the unit for Shannon variability will be bits.
-It will indicate the average number of yes/no questions we need to identify an element within the dsitribution.
+It will indicate the average number of yes/no questions we need to identify an element within the distribution.
 In general, you can pick any base $b$: in base three we have ternary questions and trits, for ten we have questions with ten possible answers and digits.
@@
 If not, we continue.
-We can group fish with reptiles and birds with small mammals to form a 13/14 split.
+We can group fish with reptiles and birds with small mammals to form a 13/12 split.
 So we can ask, ``Is it a fish or a reptile?".
@@
 It is also true that scientific theories, in the end, are models that can capture only the aspects of nature that can be tested experimentally, the information extracted by the experiment, under suitable circumstances.
-Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value'', ``the information thus solicited [by the experiment] makes physics and comes in bits'' by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations avaialble (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context in order to be understood: we need to know what the subject of our experiment is, how to prepare it and how to collect the data.
+Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value'', ``the information thus solicited [by the experiment] makes physics and comes in bits'' by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations available (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context in order to be understood: we need to know what the subject of our experiment is, how to prepare it and how to collect the data.
 The art of experimental science is contained neither in the mathematical description nor in the data collected.
@@
 This means that infinitesimally narrow distributions, like delta functions, do not work well with the Shannon variability as they would give minus infinite entropy.
-If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which have a very small but still finite support (i.e.~region where the function is non-zero), which is what we will assume.
+TODO If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which have a very small but still finite support (i.e.~region where the function is non-zero), which is what we will assume.
 
@@
 Therefore the areas $d\hat{q}^i d\hat{p}_i = dq^j dp_j$ remain the same.
-The Jacobian determinants are unitary and the Shannon variability is invariant.
+This has a couple of important consequences.
+First of all, (1) the volumes are preserved during the transformation, which means (2) the Jacobian is unitary and (3) the Shannon variability is invariant.
+In fact, these three properties are mathematically the same property: each one implies all others.
+Second, something similar happens to the areas for each independent degree of freedom, meaning that the Shannon variability associated to the marginal distributions is also invariant.
+These properties can be taken to be the defining characteristics of phase space: phase space has exactly such geometrical properties.\cite{AoP2020HamiltonianEntropy}
 
@@
 In other words, the Boltzmann entropy reduces to $N$ times the Shannon variability of the single-particle distribution for a given microstate.
+
+% http://przyrbwn.icm.edu.pl/APP/PDF/120/a120z6p04.pdf
 

### 2021-01-04 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Edits to parts of Section 4 (on new technical material, and fixed problems noted on 12-21) I still need to massively restructure this section, along with moving the discussion of the gauge redundancy of p out of section 5.2 and into this section.

--- 2020-12-21 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 In general are coordinate dependent and you can't simply compare values at different points.
-This is the case in Lagrangian mechanics because both differentials in position vary controvariantly.
+This is the case in Lagrangian mechanics because both differentials in position vary contravariantly.
 
@@
 
+\subsection{Density and Riemmanian geometry}
+
+It is helpful to take a quick look at how densities work in general so that we can better understand what is so special about them in the context in Hamiltonian mechanics.
+By density we mean the intuitive physical notion of a quantity divided by the region it occupies, like mass per volume or charge per area.
+Over a continuum, this is an integrable function $\rho(x)$.
+Mathematically, the concept is captured by the Radon–Nikodym derivative: we have a region $U$, we have two measures $m$ and $V$ and the derivative satistifies $m(U) = \int_U \rho dV$.
+This makes it clear that the units of $\rho$ depends on the units of both $m$ and $V$.
+More precisely, if $\rho(x, y)$ is a function of multiple variables, we will want $\int_U \rho(x, y) dx dy$ so that the units of $\rho$ are consistent with the units of the arguments.
+For example, in polar coordinates $\rho(r, \theta)$ will return mass over meter-radians if we measure $r$ in meters and $\theta$ in radians.
+
+The density is coordinate dependent because its definition requires us to make a choice of coordinates.
+Worse than that, it is not even coordinate invariant: the density changes with the Jacobian.
+For example, changing from polar to Cartesian coordinates we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
+Since the Jacobian is a function of the point, this makes the value of the density, in general, not directly comparable at different points.
+If we have a uniform density in $x$ and $y$, we will have an $r$-dependence in $r$ and $\theta$.
+This dependence is introduced precisely because the area identified by $dr$ and $d\theta$ is $dr \,r d\theta$, which is a function of $r$.
+
+In differential geometry, a different mathematical object is called the `density,' namely a volume form (i.e.
+an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation.
+Our physically significant density $\rho$ is actually the component of this volume form $\nu = \rho(x^i) dx^1 \wedge dx^2 \wedge ...
+\wedge dx^n$.
+Note that $\nu$ combines the density and the volume into a coordinate-free object.
+Considering the total mass in the region $U$, we have $m(U) = \int \nu$ so essentially $\nu = dm$ which is the infinitesimal mass contribution we are integrating.
+But this means the value returned by $\nu$ depends not only on how must stuff there is at a given point, but also on the size of the volume identified by the $n$ vectors given to the form.
+[Not immediately clear to me what the upshot is of this last point; e.g.
+how does this relate to $\nu$ being coordinate-free?]
+
+In Riemannian geometry, we have a canonical volume form $dV = \epsilon \sqrt{|g_{ab}|} dx^1 \wedge dx^2 \wedge ...
+\wedge dx^n$ where $|g_{ab}|$ is the determinant of the metric tensor as expressed in local coordinates, and $\epsilon$ is $1$ or $-1$ depending on the orientation.
+In component notation, we can express the volume form as $\sqrt{|g_{ab}|}\epsilon_{a_1 ...
+a_n}$, where $\epsilon_{a_1 ...
+a_n}$ is the Levi-Civita symbol.
+We can then express any differential density $\nu = \hat{\rho} dV$ as the product of the volume density times a scalar function $\hat{\rho}$.
+At first glance, this expression misleadingly suggests that the density $\nu$ is coordinate-independent, but this is not the case.
+[double check that this claim applies to $\nu$ rather than $\hat{\rho} $.
+Is this in tension with saying earlier that $\nu$ is coordinate-free?] The quantity $|g_{ab}|$ depends on the choice of coordinates and is, in general, a function of the point.
+However, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
+The determinant $|g_{ab}|$ is the Jacobian of the transformation to those coordinates [the rationale/point of this sentence was not clear to me].
+For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$; in polar coordinates $|g_{ab}| = r^2$ and the volume form $dV = r dr \wedge d\theta$ [above we said that the Jacobian was $r$ rather than $r^2$].
+This means that $\hat{\rho}$ is really expressing the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise.
+We can also see this if we look at the units of $\hat{\rho}$, which are in terms of inverse volume as expressed in the privileged system.\footnote{This is similar to what happens with the concept of proper time: it is defined in the comoving frame, so it is coordinate dependent.
+Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate invariant.}
+
+We can now look at what happens for distributions over position and velocity.
+Note that the velocities transform in the same way (i.e.
+contravariantly) as differentials in position.
+Therefore $\rho(x^i, v^j) = \sqrt{|g_{ab}|} \sqrt{|g_{ab}|} \rho(\hat{x}^i, \hat{v}^j) = |g_{ab}| \rho(\hat{x}^i, \hat{v}^j)$ $dx^a$.
+The determinant of the metric tensor is therefore directly the volume of the tangent bundle [I didn't follow this point].
+The relationship to the density as expressed in phase space is therefore the following:
+\begin{equation}
+	\rho(q^i, p_j) = |g_{ab}| \rho(x^i, v^j).
+\end{equation}
+A change of coordinates will induce a contravariant change in the velocity and a covariant change in the momentum.
+The left side will remain invariant.
+The two terms on the right side will change in the opposite way so that their product is invariant.
+Focusing on densities, then, brings about a physically-motivated way to understand the relationship between the cotangent and the tangent bundle, where the metric tensor plays a role in defining the geometry of the second in a similar way that the symplectic form defines the geometry of the first: it defines the count of states so that densities can be properly tracked.
+
 ---
@@
 Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves.
--> Invariant density is really the important physical element/object in Hamiltonian mechanics.
+$->$ Invariant density is really the important physical element/object in Hamiltonian mechanics.
 
@@
 
-* Invariant densities: physical significance connected to coordinate independence. [incorporated]
+* Invariant densities: physical significance connected to coordinate invariance. [incorporated]
 
@@
 They are coordinate invariant.
-Also the Shannon/Gibbs entropy are coordinate independent (see physicality of quantities that are coordinate invariant)
+Also the Shannon/Gibbs entropy are coordinate invariant [but not coordinate independent?] (see physicality of quantities that are coordinate invariant)
 
@@
 In general, it provides a better understanding of classical systems.
-Specifically, we will see that invariant densities are of fundamental physical significance due to their coordinate-independence [invariance?].
+Specifically, we will see that densities are of fundamental physical significance due to their coordinate-invariance.
 Hamiltonian mechanics---unlike Lagrangian mechanics---allows us to compare these densities point-wise in any coordinate system.
-Although the density is physically significant because it is coordinate-invariant, we will nevertheless argue that coordinate-invariance is not a necessary condition for physical significance.
-In fact, the ability to express densities in some choice of units is closely connected with symplectic structure.
+Although the density is coordinate-invariant, it is not coordinate-independent (it cannot be given a coordinate-free definition).
+This is because expressing the density requires a choice of units, a fact that is closely connected with symplectic structure.
 Hence, we will argue that some coordinate-dependent representational choices are nevertheless physically significant.
@@
 % Section~\ref{action} uses the invariant density to provide a geometrical interpretation of the action principle.
+% Remember: the density is coordinate-invariant but not coordinate-independent; it is actually coordinate-dependent (i.e. it does not admit of something like a coordinate-free definition)
+% Also remember: there are quantities that are both coordinate-invariant and coordinate -dependent that are still physically significant, e.g. choices of q's and p's (?)
+% Coordinate-invariant in general means that definition of the quantity involves referencing coordinates, but if you change the coordinates, then the value of the quantity remains the same.
+(E.g.
+transforms as a scalar).
+so densities are never coordinate-independent.
+defining a density requires a choice of units, so involves referencing units.
+%see 57.3 minutes of 12-21 recording: when we define the density originally, it is coordinate-free (rho as a function of the state space of the particles, S).
+But when we go to express it numerically, then it becomes coordinate-dependent.
+(I.e.
+when we express it over a manifold, then it needs to be a density for integration, and hence it becomes coordinate-dependent).
+% see 51 minutes of 12-21 recording: Densities (the coordinate-dependent kind) are physically significant partly because they tell you how much stuff is in each state
 
@@
 For brevity, will refer to this as the property of \textit{point-wise comparison}.
-Since the density depends in general on the choice of units, point-wise comparison does not hold universally across representations of state space.
-Instead, for the density of states to have this property, we require a coordinate-independent definition of the density.
+Since in general the density is coordinate-dependent, this property of point-wise comparison does not always hold.
+For instance, it can fail as we change from Cartesian to polar coordinates [figure out where to insert or reference the illustration of this shown above].
+For point-wise comparison to hold, we require a coordinate-invariant density.
 As the derivation in Section~\ref{deterministic} shows, we must define our state space using a symplectic manifold in order to define an invariant density.
@@
 Hence, the real physical significance of canonical coordinates is that they allow us to define an invariant phase space density $\rho $.
+[Move this section 5.2 stuff up into this section; and figure out how to reorganize this paragraph]
 
@@
 The first is epistemic: by representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
-Any and all coordinate systems will support the formal property of point-wise comparisons [what about time-dependent coordinate transformations?].
-Not only is this formal property methodologically advantageous, but it also bespeaks a further physical significance.
-This brings us to a second, physical notion of fundamentality: the phase space density $\rho$ is physically fundamental because it is a coordinate-invariant.
-Assuming that coordinate-invariance is sufficient for physical significance, the density thereby represents a genuine physical property of classical systems.
+Any and all (time-independent) coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
+See~\ref{gauge} for discussion.} Not only is this formal property methodologically advantageous, but it also bespeaks a further physical significance.
+This brings us to a second, physical notion of fundamentality: the phase space density $\rho$ is physically fundamental because it tracks how much of the system is in each possible state.
 It quantifies the amount of stuff in each phase space state.
-[to do: should we make this precise in terms of representational theory of measurement?
-E.g., the particular kind of scale that this representation defines and the meaningful comparisons it supports?].
-
-% Here, we consider an epistemic sense of ``fundamental." In this sense, Hamiltonian mechanics is ``fundamental" because it shows that densities are coordinate-independent quantities.
-Assuming that coordinate-independence is sufficient for being physically significant, this shows that these densities are physically significant.
-More precisely, they are physical quantities or properties of the system.
+Section~\ref{deterministic} demonstrated what is perhaps a third reason for viewing the density as fundamental: by defining the density, we recover the right mathematical structure, namely a symplectic manifold.
+If we did not define the density, then we would need another explanation of why this mathematical structure arises for classical systems.
+[I think this final point might be repetitive given a claim made in the next paragraph]
+%see 11:50am of 12-21 meeting video for clarification of this point. There are probably finer details for me to incorporate.
+
+% We took out the following because coordinate-invariance in fact did not seem to be either a necessary or sufficient condition for physical significance: Assuming that coordinate-invariance is sufficient for physical significance, the density thereby represents a genuine physical property of classical systems.
 

### 2021-01-04 Articles/2019-ShannonEntropy/Submissions/EJP/response.tex
Preparing response for EJP

@@
 
-TODO add footnote
+We added a footnote that reads:
+
+\emph{We will not discuss its relation to thermodynamic entropy.
+Thermodynamic entropy is defined on the notions of heat and work while the Shannon entropy is defined on the notion of a distribution.
+The contexts are different and relating them goes beyond the scope of this work.}
 
@@
 
-TODO: add here that similar comments are throughout the paper
+There are multiple comments to that effect throughout the paper.
+The referee is attributing us an ``implicit'' thesis that is not our own, instead of confronting the explicit ``thesis'' that is repeated throughout the paper.
+It is not apparent to us how we could modify the manuscript to avoid that, particularly because he is the only one that has done so.
 
@@
 \end{quote}
-That reference appears after the sentence ``it is still surrounded by a great deal of confusion''.
+That reference appeared after the sentence ``it is still surrounded by a great deal of confusion''.
 Therefore it is only used as evidence to show that there is confusion surrounding entropy.
@@
 
-The referee, once again, puts words in our mouth that are not on the page. The use of `variability' for the Shannon formula is justified by this paragraph:
+The referee puts words in our mouth that are not on the page. The use of `variability' for the Shannon formula is justified by this paragraph:
 
@@
 ``Entropy'' is used for the physical concepts in statistical mechanics.
-That helps the reader understand that some properties of that concept are more general than physics.
+That helps the reader understand that some properties of that concept are more general than physics, which is the proper setting to understand the concept.
 Given that we make a distinction between the two words, ``variability'' and ``entropy'' are \textbf{not} synonyms in our work.
 That is the whole point.
-The variability of species in ecology has nothing to do with entropy in statistical mechanics.
+The variability of species in ecology is not a measure of entropy in statistical mechanics.
 
@@
 \end{quote}
-This claim is hotly contested in the literature of the foundations of thermodynamics.
-[cite] In modern textbooks of statistical mechanics, the Shannon entropy is not the starting point, rather an approximation that is valid only at equilibrium [cite].
-The referee is therefore expressing his personal view.
+The referee is expressing a view that, though shared by others, is contested in the literature of the foundations of thermodynamics.
+For example, see ``Time-reversal and entropy'' by Christian Maes et al, or ``Gibbs and Boltzmann Entropy in Classical and Quantum Mechanics''
+Sheldon Goldstein et al.
+In many modern textbooks of statistical mechanics, the Shannon entropy is not the starting point, rather an approximation that is valid only at equilibrium.
+See, for example, ``Statistical mechanics in a nutshell'' by Luca Peliti, where the starting point is state counting.
+We are staying out of that debate as we would find inappropriate, in the context of a paper aimed at scientific education, to push a particular view.
 
-We are staying out of that debate and we would find inappropriate, in the context of a paper aimed at scientific education, to take a side as the referee suggests.
+We do believe that part of the confusion surrounding entropy does come from these unresolved contentions.
+This was clear while doing literature review with some of our students.
+We therefore added the following sentence in the introduction:
 
-TODO: add in the introduction
+\emph{[...
+entropy ...] is surrounded by a great deal of confusion.
+Partly because, at a fundamental level, there is still disagreement on what is the ``correct'' notion of entropy, particularly out of equilibrium.}
 
@@
 
-This is beyond the scope of the paper, which is simply to introduce in a clear way to newcomers what the expression $- \sum p \, \log p$ represents.
+Doing a comparative analysis of this kind is beyond the scope of the paper, which is simply to introduce in a clear way to newcomers what the expression $- \sum p \, \log p$ represents.
+See, for example, "Compendium of the foundations of classical statistical physics" by Jos Uffink.
 
@@
 \end{quote}
-Again, we are in no way disproving what he says and the referee is engaging in what philosophers define as ``verbal dispute'', where one disagrees on the meaning of the words rather than on an actual fact of the matter.
-We mean `subjective' in the Bayesian/machine learning sense, where each agent has a different take on how much something is likely.
+Again, we are in no way disproving what Jaynes says.
+What the referee (but he is not the only one) is engaging in what philosophers define as ``verbal dispute'', where one disagrees on the meaning of the words rather than on an actual fact of the matter.
+The problem is what one means by ``subjective''.
 Jaynes, unfortunately, used language that sometimes seems Bayesian to refer to cases where the experimental setup matters (which is not Bayesian).
@@
 
-Given that the referee keeps mis-characterizing our prose, it is difficult for us to understand what, if anything, gave him the wrong impression.
-Note that we tested the article on a group of students and none of these misunderstandings emerged.
-
-TODO: review the 1992 paper, see if there is something about physicality
+We have expanded the section to clarify that and also cited the 1992 which does not change anything. [TODO include]
 
@@
 The resolution does not fix the possible negative value of the Shannon entropy for continuous distributions.
-The importance of granularity, of unit choice, not only was already present in the introduction, but is literally given in the following paragraph of the original submission:
+The importance of granularity, of unit choice, not only was already present in the introduction, but is literally given in the original submission in paragraph following the one the referee refers to:
 
@@
 
-Again, given the complete mis-characterization of the prose and the erroneous claim that somehow the choice of unit/granularity makes the Shannon entropy always positive, we find it impossible to understand whether there is any meaningful item to address.
-
-TODO: add "granularity"
+Given the complete mis-characterization of the prose and the erroneous claim that somehow the choice of unit/granularity makes the Shannon entropy always positive, we find it impossible to understand whether there is any meaningful item to address.
+One possible answer is that the referreee does not realize that choosing a unit, in the context of information entropy, is equivalent to choosing the resolution since it changes the zero for entropy...
+The $\Delta$ \textbf{is} the unit.
+We added the words ``granularity'' and ``resolution'' in a few places where units are discussed.
 
@@
 
-TODO: add footnote
+We added the following footnotes, merely as a pointer for the curious reader. [TODO]
 
@@
 \end{quote}
+
+This referee report, while abundant in ad hominen, mischaracterizations and demeaning language, was outstandingly lacking in constructive criticism.
+This is peer reviewing at its worst.
+The most likely scenario is that, after a very cursory look, he framed the manuscript in some diatribe the referee has with some other colleagues, stopped reading carefully and engaged in a tirade, assigning to us views that are not our own.
+This means that most of the comments weren't even directly relevant to our manuscript: even after sifting the report carefully, we found few actionable items.
+
+We urge the referee to conduct himself in a more professional manner.
+Whether or not variability is a better characterization for the Shannon entropy can be debated rationally without resorting to mockery.
+
+\section*{Referee 3}
+
+\begin{quote}
+	I think that the manuscript is well written, well organised and generally speaking, the main topic is interesting also for a not expert reader. 
+\end{quote}
+
+We thank the referee for the kind words.
+
+\begin{quote}
+The idea of the Shannon entropy as a way to quantify the number of yes/no questions (i.e.
+bits) that are needed to identify an element within the distribution is not new.
+However, I appreciate the authors' presentation because this could have validity in the field of Physics Education.
+\end{quote}
+
+This is our belief as well and one of the reasons we wrote the article.
+
+\begin{quote}
+Although I think that the manuscript deserves to be published, I would like that the authors add some bibliographic references on the topic they deal with but in the field of Physics Education.
+\end{quote}
+
+Most of the literature we found in Physics Education focuses on the Gibbs, Boltzmann and Thermodynamic entropies.
+We added references for those in the first paragraph of the introduction, along with a remark stating that the Shannon entropy itself has received less attention.
+
+\begin{quote}
+As a minor issue, I noted a typo at line 27.
+\end{quote}
+
+\section*{Referee 4}
+
+\begin{quote}
+After reading the paper, Referee’s and Adjudicator’s comments I think that it is proper to send all the correspondence to the Authors.
+I suggest the Authors to give their replies directly on the pages of the paper: such and similar comments/questions appear in a natural way when one reads the manuscript.
+The paper (and readers as well) will benefit from modifications along the lines raised by Referees.
+In fact, I am convinced that  Referee 2 reflections  (the most critical one) will be very useful in this context and can be properly incorporated in the paper.
+I think that they are well based and clearly explained, however, I do not think that the Authors ‘...have created an ...arguments against the Shannon entropy..’ (as the Referee states).
+Rather they aim to explain for a wide audience that ‘the Shannon entropy measures the variability of the elements within a given distribution, giving it a crisp intuitive meaning that is general and applicable to all branches of science’ (as the Authors say in their Conclusions).
+\end{quote}
+
+We thank the referee for appreciating the goal and the intended audience of the paper. We tried addressing as best we could the points raised by the second referee.
 

### 2021-01-04 Articles/2019-ShannonEntropy/Submissions/EJP/EJP-ShannonEntropyIsVariability.tex
Preparing response for EJP

@@
 The general concept of entropy in physics, despite being treated extensively in the physics education literature (e.g.
-[cite physed]), is still surrounded by a great deal of confusion [cite evidence of confusion]\cite{Swendsen, KishFerry}.Partly because, at a fundamental level, there is still disagreement on what is the ``correct'' notion of entropy, particularly out of equilibrium.
-[cite works of out of equilibrium research]  We will focus on the notion of Shannon entropy, which has received less attention in physics education.
-We will see that the Shannon entropy can be given a crisp and precise characterization and will explore its relation to the Gibbs and Boltzmann entropies in statistical mechanics.\footnote{We will not discuss its relation to thermodynamic entropy: this requires the notions of heat and temperature, which will not be covered in this work.}
+[TODO cite physed]), is still surrounded by a great deal of confusion [TODO cite evidence of confusion]\cite{Swendsen, KishFerry}.
+Partly because, at a fundamental level, there is still disagreement on what is the ``correct'' notion of entropy, particularly out of equilibrium.
+[TODO cite works of out of equilibrium research] Partly because the same word ``entropy'' is used to refer to concepts (e.g.
+topological entropy, thermodynamic entropy, graph entropy, R'{e}nyi entropy, information entropy) that have different definition and various degree of overlap.
+We will focus on the notion of Shannon entropy, by which we mean the use of $- \sum_i p_i \log p_i$ in the broadest characterization, which has received less attention in physics education.
+We will see that the Shannon entropy can be given a crisp and precise characterization and will explore its relation to the Gibbs and Boltzmann entropies in statistical mechanics.\footnote{We will not discuss its relation to thermodynamic entropy.
+Thermodynamic entropy is defined on the notions of heat and work while the Shannon entropy is defined on the notion of a distribution.
+The contexts are different and relating them goes beyond the scope of this work.
+}
 
@@
 A distribution over half a unit interval will correspond to minus one bit, since we would need to ``lose'' one bit of information to widen the variability back to one unit.
-In other words, the variability is measured by the number of bits needed to identify an element up to one unit.
+In other words, the variability is measured by the number of bits needed to identify an element up to one unit, up to the resolution (i.e.
+the granularity) given by that unit.
 
 This means that infinitesimally narrow distributions, like delta functions, do not work well with the Shannon variability as they would give minus infinite entropy.
-TODO If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which have a very small but still finite support (i.e.~region where the function is non-zero), which is what we will assume.
+If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which  maybe have very small but still finite support (i.e.~region where the function is non-zero), which is what we will assume.
 
@@
 
-The Boltzmann constant $k_B$ should not distract us: its role is simply to allow us to measure temperature in an appropriate unit.
+The Boltzmann constant $k_B$ should not distract us: its role is simply to allow us to measure temperature in an appropriate unit.\footnote{In the latest revision of the International Metric System of Units SI,\cite{chyla2011evolution} the Boltzmann constant is one of the defining constants.
+This means its value is assigned, rather then measured, in a way that properly defines Kelvin, the unit of temperature.
+The universal constants are essentially used as fixed starting points to define our measurement scales.
+}
 If one defines $T = \frac{\partial U}{\partial S}$, and measures temperature in Kelvin and energy in Joules, the above relationship forces us to measure entropy in Joules/Kelvin.
@@
 However, this is not the only possible definition.
-Instead of using $T$ as a primary thermodynamic variable, we can use $\beta = \frac{1}{k_B T}$.\footnote{There are a few reasons why one can argue this is a better choice, though ultimately it is still a matter of choice.} In this case, one would define $\beta = \frac{\partial S}{\partial U}$, measure energy in Joules, $\beta$ in inverse Joules and entropy would be dimensionless.
-As one can do all the thermodynamic calculations using just $\beta$, the constant $k_B$ is really just set by the unit system.
+Instead of using $T$ as a primary thermodynamic variable, we can use $\beta = \frac{1}{k_B T}$.
+In this case, one would define $\beta = \frac{\partial S}{\partial U}$, measure energy in Joules, $\beta$ in inverse Joules and entropy would be dimensionless.
+As one can do all the thermodynamic calculations using just $\beta$, the constant $k_B$ is really just set by the unit system.\footnote{As \cite{chyla2011evolution} notes:
+\begin{quote}
+	Temperature characterizes the average thermal energy of particles in a certain ensemble of particles in the state of thermodynamic equilibrium.
+Therefore, in principle, one could express this quantity in terms of the unit of energy; this is actually practiced in many fields of physics, especially in statistical physics, where it is quite common to use $\Theta = k_B T$ or $\beta = 1 / k_B T$  as a measure of thermal energy, instead of temperature $T$ measured in Kelvins.
+\end{quote}
+}
 
@@
 We note that some authors choose to interpret the probability distribution not as coming from repeated independent trials, but as the knowledge one has about the system.
-This would make the entropy a subjective notion\footnote{Jaynes\cite{Jaynes} called it the anthropomorphic nature of entropy.}, but the associated fluctuations we experimentally observe are indeed objective.
-We believe the confusion comes from the, correct, realization that the same system, under different conditions, will be described with a different set of thermodynamic variables.\footnote{Jaynes\cite{Jaynes} points out:
+This would make the entropy a subjective notion: each observer would have a difference credence distribution, regardless of whether it fits the data, and therefore a different Shannon entropy.
+A physical quantity, however, must be the same for everybody and the associated fluctuations we experimentally observe are indeed objective.
+So what is going on?
+
+We believe the confusion comes from the, correct, realization that the same system, under different conditions, will be described with a different set of thermodynamic variables.
+\footnote{Jaynes\cite{Jaynes} points out:
 \begin{quote}
@@
 This is what we stated at the beginning: a distribution, and therefore its variability, is always contingent upon some arbitrary choices.
-The question, then, is whether the system plus the process plus the set of constraints are enough to determine the entropy, or if we also need to mix in human knowledge.
-What we find is that they are sufficient, so the entropy does not depend on human knowledge.
+Since the Gibbs entropy is the variability of a microstate within the distribution identified by the macrostate, it is not a property of the single microstate, it is not a property of the system, but it is a property of the system within that specific process, of the macrostate.
+Some authors refer to this fact by saying that entropy is not objective \footnote{Jaynes\cite{Jaynes,Jaynes2} called it the anthropomorphic nature of entropy.}.
+
+The issue is that subjective, in the context of probability theory, refers to Bayesian probability, which really means subjective: what ones believes to be true regardless of what experimental evidence there is.
+This is not at all the same concept as the one outlined before.\footnote{Note how Jaynes\cite{Jaynes2} always put ``subjective'' in double quotes.} The system plus the process (which identifies the set of constraints) determines the ensemble and therefore the entropy, regardless of whether an agent knows what process was used.
+Even if one wants to give a Bayesian account, then, one has to give it in terms of an agent that has full and exact knowledge of the system and the process.
+But this is hardly subjective.
+In our view, this is a case of unfortunate word choice.
+It is much better to simply state that the entropy is a property of the ensemble, of the distribution, and not of the system.
+And if understand that the Gibbs entropy is the variability of the microstate within the ensemble, then it is clear that it cannot be a property of the microstate itself, but it is a property of the ensemble.
 

### 2021-01-05 Articles/2019-Entropy/Source/thermodynamics.tex
First draft for thermodynamics

@@
 \title{Counting evolutions: a simple foundation for thermodynamics}
-\author{Gabriele Carcassi, Christine A. Aidala, et al \\ University of Michigan}
+\author{Gabriele Carcassi\footnote{carcassi@umich.edu}, Christine A. Aidala, et al \\ University of Michigan}
 
@@
 
-This draft summarizes the status of our work on a new approach to the foundations of thermodynamics and is intended to gather early feedback.
+This working draft summarizes the status of our work on a new approach to the foundations of thermodynamics and is intended to gather early feedback.
 Rather then derive thermodynamics bottom-up from a particular microscopic model, we want to derive it top-down from a set of requirements on the dynamics of the system.
 The results we find are, then, independent on the microscopic dynamics in the sense that they can in principle be realized in different ways.
-This is both more in line with the aims of the founders of thermodynamics and justifies why similar concepts have been successful in fields beyond physics.
-Moreover, thermodynamics is found as a specialization of a more general model, which we believe can be used as a common foundation for the different types mechanics.
+This is both more in line with the aims of the founders of thermodynamics and justifies why thermodynamic concepts have been successful in fields beyond physics.
+Moreover, in our approach thermodynamics is found as a specialization of a more general model, which we believe can be used as a common foundation for the different types mechanics.
 
@@
 However, those do not really change the overall intuition, which we believe is the main strength of the approach.
-Therefore, in this paper, we will only give a conceptual overview leaving the formal details to other.
-This will leave the key physical insight is more directly accessible, and also result in a narrative that can be used directly in an introductory course in thermodynamics.
+Therefore, in this paper, we will only give a conceptual overview leaving the formal details to other papers.
+This will leave the key physical insight is more directly accessible, and also result in a narrative that can be used directly in an introductory course of thermodynamics.
 
@@
 
-Conversely, we say a process is \textbf{reversible} if the state at one time is enough to reconstruct the state at all past times.
-We can see an example of such a process in figure \ref{fig_reversibility}.
+Conversely, we say a process is \textbf{reversible} if the state at one time is enough to reconstruct the state at all past times.\footnote{Note that in some literature the ability to reconstruct the past is called retrodictability and does not coincide with thermodynamic reversibility (the ability to undo a change).
+This distinction appears in the context of continuous quantities and it is precisely one of those technical considerations we (will) address in a more technical work.
+At this point we want to make it clear that our notion of reversibility will indeed correctly recover thermodynamic reversibility.
+It will do so without having to assume the existence of a reverse process or a symmetry under time-reversal.} We can see an example of such a process in figure \ref{fig_reversibility}.
 In this case, if two different evolutions happen to reach the same state at a given time, they will also share all past states: evolutions can diverge but cannot merge.
@@
 
-Note that in some literature the ability to reconstruct the past is called retrodictability and does not coincide with thermodynamic reversibility (the ability to undo a change).
-This distinction will be made precise and clear in the full framework.
-At this point we want to make it clear that our notion of reversibility will indeed correctly recover thermodynamic reversibility.
-It will do so without having to assume the existence of a reverse process or a symmetry under time-reversal.
-
 A process that is both deterministic and reversible, like the one in figure \ref{fig_detrev}, will never contain evolutions that split or merge.
@@
 
+%TODO: add a double evolution in some case.
 \begin{figure}[h!]
 	\includegraphics[width=\columnwidth]{images/Slide5.png}
-	\caption{Determinism and reversibility. Evolutions cannot merge nor split. The number of evolutions per state cannot increase.}\label{fig_detrev}
+	\caption{Determinism and reversibility. Evolutions cannot merge nor split. The number of evolutions per state cannot increase.} \label{fig_detrev}
 \end{figure}
 
-To finish, we focus on deterministic processes with equilibria.
+To finish, consider a \textbf{deterministic processes with equilibria}.
 We can see an example of such a process in figure \ref{fig_with_equilibria}.
-An equilibrium for a process is a state that, once it is reached, it is never left.
-In a process with equilibria all evolutions will reach, at some point, an equilibrium.
+An equilibrium for a process is a state that, once it is reached, it is never left.\footnote{We will see later that this characterization of equilibrium is symplistic, but it will suffice for now.} In a process with equilibria all evolutions will reach, at some point, an equilibrium.
 In the out-of-equilibrium phase, evolutions merge until they reach equilibria and can no longer merge.
@@
 If we care to characterize only the initial and final state, deterministic evolution simply means that the final equilibrium is a function of the initial state.
-
-Note that the idea of equilibrium as ``nothing changing'' used here is an oversimplification, which does not even work relativistically.
-Yet, it suffices for now as it captures the intuition correctly.
 
@@
 
-Note that we have found a quantity, $\mu(x(t))$, that can never decrease in a deterministic process, stays the same if the process is reversible and is maximized at equilibrium.
+We already reached an important result: with only a handful of simple definitions we have found a quantity, $\mu(x(t))$, that can never decrease in a deterministic process, stays the same if the process is reversible and is maximized at equilibrium.
 We have done so with no assumption on the type of system, the type of forces acting on the system, the internal composition of the system or the internal dynamics of the system.
@@
 The result is completely general and the justification is very straight forward.
-It is conceptually a perfect candidate for the thermodynamic entropy, expect for one thing we still need to understand: how it combines under system composition.
+It is conceptually a perfect candidate for the thermodynamic entropy, though we have one questions left: how does it combine under system composition?
 
@@
 A process defined on the composite system will come with a measure $\mu_{X \times Y}$ that enables us to count evolutions $\{x(t), y(t)\}$ of the composite system.
-In general, the state of one system may affect the other system, in which case the evolution of the composite cannot be separated.
-If the systems are independent, however, the evolution of one system will not influence the other.
+In general, the evolution of one system will affect the evolution of the other system.
+This will not happen, however, if the systems are independent.
 That is, for each possible evolution $x(t)$ of the first system and $y(t)$ of the second system, there will be a possible evolution $\{x(t), y(t)\}$ on the composite.
@@
 \end{equation}
-the measure factorizes.
+the measure factorizes. We say two system are \textbf{independent} at given time $t$ if the evolutions per state factorizes: $\mu_{X \times Y}(x(t), y(t)) = \mu_X (x(t)) \mu_Y (y(t))$.
+
 We define the \textbf{process entropy} as $S = \log(\mu)$.
@@
 \end{equation}
-In addition, the process entropy will be additive for independent systems.
-As before there are a number of technical details that need to be addressed, most of all the units for $\mu$ which will influence the 0 point for $S$, which we will address in the full framework.
+In addition, the process entropy will be additive for independent systems.\footnote{As before there are a number of technical details that need to be addressed, most of all the units for $\mu$ which will influence the 0 point for $S$.
+We (will) address in the full framework.}
 
-In general, the process entropy will depend on the process, not just on the state.
-However, a unique value for process entropy can be assigned to states of equilibrium.
+Under very minimal conditions, we have found a quantity that has all the right properties for entropy and that is linked with determinism and irreversibility in a plain, intuitive and physically meaningful way.
+
+\section{System independence and state entropy}
+
+The careful reader will have noticed that the process entropy will depend on the process, not just on the state.
+That is, the same description of the same system may be associated with more or fewer evolutions depending on the conditions.
+Entropy in physics, however, is associated to a state.
+Under what conditions can we associate a unique process entropy to a state?
+
+A unique value for process entropy can be assigned to states of equilibrium.
 Suppose, in fact, that the state $x$ is an equilibrium for two different processes $A$ and $B$.
-Suppose that $A$ is at equilibrium $x$ and we switch to process $B$.
+Suppose that at the end of process $A$ the system is at equilibrium $x$ and we switch to process $B$.
 Since $x$ is also an equilibrium for $B$, every evolution of $A$ that ended in $x$ will correspond to an evolution in $B$ that ends in $x$.
@@
 Therefore the process entropy of an equilibrium does not depend on the process, only on the equilibrium state.
-This allows us to assign a unique process entropy to the state.
-If we have set of states that are equilibria, then, we call \textbf{state entropy} $S(x)$ the process entropy that the system reaches at equilibrium.
 
-To summarize, we found, under most general conditions, a quantity called process entropy, defined as the logarithm of the number of evolutions, that cannot decrease under deterministic processes, cannot increase under reverersible processes, is maximized at equilibrium and is additive for independent systems.
-Moreover, if $X$ consists of equilibria, there exists an \textbf{equation of state} $S : X \to \mathbb{R}$ which returns the state entropy, which is the process entropy of the equilibrium.
+If the system is a composite there are additional considerations.
+Different processes may put different degree of correlation between subsystems, therefore the same equilibria for the subsystems may correspond to different processes entropy values for the composite.
+But if the subsystems are independent, the composite will correspond to a unique entropy value: the product.
+
+We can then assign a unique process entropy to states of equilibria and to composite state of independent equilibria.
+In these conditions we define the \textbf{state entropy} $S(x)$ which corresponds to the process entropy that the system reaches at equilibrium.
+But why would we want to assign a unique process entropy to states in the first place?
+Why would this be interesting?
+There are three main reasons.
+
+The first is process combination.
+Suppose that we are studying a system and we change the circumstances surrounding it.
+It first evolves according to $A$ and we switch to process $B$.
+If $x(t)$ is an evolution of the combined process $AB$, then for $t\leq t_0$ it must be an evolution for $A$ and for $t\geq t_0$ it must be an evolution for $B$.
+Moreover, the number of total evolutions must remain the same.
+Therefore the state $x(t_0)$, the state where we are joining the evolutions, must have a well defined entropy that is independent on the process, or not all the evolutions would match.
+If we want to mix and match from a set of possible processes, then the process entropy at the junction must be only a function of the state.
+
+The second reason is that, upon close inspection, measurements are never defined at a instant of time.
+Therefore $x(t)$, the state at an instant of time, is really a shorthand for $x[(t - dt/2, t + dt/2)]$, the state defined over a small open interval $dt$ around $t$.
+We can talk about a single state within that interval only if all state variables do not change and the system is fully characterized by those state variables alone.
+Therefore writing $x(t)$ implicitly assumes two things: that the system is an equilibrium of the processes that happen at a faster scale of our measurements of said system; that further description (e.g.
+the internal dynamics of the constituents at a finer scale) is not needed to describe the process.
+
+The third reason is that by state we implicitly mean a description of a system and only about that system.
+If there are correlations between that system and others at a time, then the description of one will tell us more than that system alone.
+Proper states are then defined in the case where the system is independent, where all the correlations have been removed.
+This is true also for correlations with the environment and correlations between subsystems.
+But the condition of independence is precisely the condition that maximizes entropy, making it uniquely defined.
+
+All three insights points toward the same directly: the very notion of state seems to be intrinsically associated to the notion of equilibrium and state entropy.
+This would suggest that all states for all systems are, in a sense, equilibria.
+But what is exactly this sense?
+What does it mean that a mechanical system, classical or quantum, is at equilibrium?
+
+Even in thermodynamics, equilibrium is usually an ill-defined concepts.
+It is typically associated the one coming from dynamical systems (i.e.
+the system no longer changes), but this doesn't quite work.
+In fact, it is not even relativistically consistent.
+If we have a volume of gas in thermodynamic equilibrium, we would like to say it is in equilibrium for any observer, not just the one in its rest frame.
+A booster observer, then, would see an equilibria where the position is indeed changing.
+In presence of gravitational forces, this motion would not even be linear.
+The notion of equilibrium as ``nothing is changing'', then, is not quite right and should be amended.
+But how?
+The previous discussion would hint toward system independence, but why system independence should be an equilibrium?
+
+The insight is that we can define something as a system only when we are able to take it apart from the rest and study it independently.
+Therefore we need, at least conceptually, a process that takes that something and makes it independent from the rest, decouples it, makes it autonomous.
+Our ability to define a system implicitly assumes the existence of such ``decoupling'' process.
+Note that if the system is already independent, the decoupling process does nothing: in this sense independent systems are an equilibria of a decoupling process.
+
+This ``decoupling'' process is what maximizes entropy.
+If the system is truly rendered independent, then we cannot know more than the description given by the state at that particular time.
+We cannot know how the system came to be in that state, we cannot know anything more about the subsystems, other systems or the environment than what the state of the system strictly tell us.
+All possible ways that the state can be realized must be realized in a possible evolution, or we could infer something more.
+When the system is independent, the states tell us the least possible information about everything else and the evolution of anything else.
+
+The definition of a particular system, then, is conditional on the existence of particular processes: those that render the system well defined.
+We can talk about the motion of a ball because we are on earth at 20 Celsius.
+If we were on the surface of the sun, we would not be able to talk about the ball in the first place.
+To define a system requires us to define a boundary, which should remain the same during the process.
+In mathematical terms, the state space must be a symmetry (i.e.
+an invariant, an equilibrium) of a group (i.e.
+a set of transformation, a set of processes).
+
+The full mathematical development of these concepts is left for another work, but the conceptual foundations are simple and direct.
+Grouping quantities into cohesive units that ``belong together'', into systems, comes with a set of practical requirements: we need a way to prepare and study them independently.
+This notion of ``independence equilibrium'' is what we will use as a foundation for ``thermodynamic equilibrium'', and is generalized to all systems.
+It is the necessary and sufficient condition to assign state entropies.
+In this light, entropy maximization is not a universal law of the universe, but a necessary requirement to build models.
+We can only study the state of objects that allow process entropy maximization.
+What remains is to answer one last fundamental question: is state entropy at all related to thermodynamic entropy?
 
@@
 
-We now have all the basic tools we need to reconstruct thermodynamics, which we characterize as the study of equilibria of composite systems.
-This is not done in general, but under additional assumptions and definitions that tell us how the equilibrum of the whole relates to the equilibria of the parts.
-As such, the laws we find are not necessarily universal in nature: they apply to systems that satisfy (or can be assumed to approximately satisfy) the prerequisites.
+The notion of equilibrium as independence is in general not sufficient to recover thermodynamic equilibrium. So, how are they related? When does the second arise?
+
+We saw how the notion of state as independence equilibrium came naturally when considering process composition of a single system.
+When the process changes, we need to reconcile the count of evolution at the junction.
+We now want to consider the scenario of process composition of composite systems.
+The idea is that we want to describe the situation in which two systems come together, evolve as a composite for some time, and then proceed to evolve separately.
+In this case, at each juncture we have to reconcile the evolutions of both the whole and each of the part.
+We need to have an independence equilibrium at both levels.
+What we'll find is that thermodynamics studies this case.
+But let's proceed in order.
 
@@
 A \textbf{thermodynamic process} is a process with equilibria where the final state is a thermodynamic equilibrium.
-We again stress that we are not saying that two systems, in general, always allow for thermodynamic equilibria.
+We stress that we are not saying that two systems, in general, always allow for thermodynamic equilibria.
+However, we contend that these are the type of processes we can mix and match to create composite processes of composite systems.
+What we want to explore is what happens if we had such systems and process, at least to a first approximation.
 
@@
 The characterization of thermodynamic equilibrium we give is manifestly a part-whole relationship.
-It extends the notion of equilibrium to a composite system so that the equilibrium of the whole and the equilibrium of its parts are essentially the same equilibrium.
-We find that this kind of part-whole relationships are a good starting point not only for thermodynamics, but also for the other branches of physics.
+It extends the notion of independence equilibrium to a composite system so that the equilibrium of the whole and the equilibrium of its parts are essentially the same equilibrium.
+We find that clarifying this kind of part-whole relationships make for a good conceptual starting point not only for thermodynamics, but also for the other branches of physics.
 This should not come as a surprise as reductionism (i.e.
@@
 
-We define a \textbf{thermodynamic system} one whose equilibria can be identified by a set of extensive quantities (i.e.
-additive under system composition).
-Moreover, one such quantity is called internal energy which has the property of remaining conserved under any process.
+Another important question for a part-whole relationships is: how are the state variables of the whole related to the state variables of the parts?
+We define a \textbf{thermodynamic system} one that follows these rules:
+\begin{itemize}
+	\item states (i.e.
+independence equilibria) are identified by a set of extensive quantities (i.e.
+additive under system composition)\footnote{It may be possible that the existence of extensive quantities can be derived from the composability of thermodynamic equilibria.
+Note, in fact, that we typically construct measurement scales of quantities (such as energy, mass, volume, distance, ...) by preparing unit samples and then adding or dividing them.
+Assuming the existence of thermodynamic equilibria may enable us similar constructions without extra assumptions.
+(XXX: any work related to this?) Nevertheless, for this work, we consider the existence of extensive quantities an extra assumption.
+	}
+	\item one such quantity is called internal energy
+	\item internal energy will be conserved under any deterministic process for that system.
+\end{itemize}
 We will denote $U$ as the internal energy and $x^i$ as the remaining extensive quantities.
+
 For a thermodynamic system, then, the state entropy can be expressed as $S(U, x^i)$ a function of the extensive quantities.
@@
 which are the usual thermodynamic relationships, except that, for us, the state entropy $S$ dimensionless..
-
-Note, again, that the system is identified by quantities that have a clear rule under system composition.
-We suspect that the existence of extensive quantities can be derived from the composability of thermodynamic equilibria.
-In fact, we typically construct measurement scales of quantities (such as energy, mass, volume, distance, ...) by preparing unit samples and then adding or dividing them.
-So it may be possible that, once we assume the ability to combine systems into bigger systems, we will consequently be able to construct extensive quantities without extra assumptions.\footnote{XXX: any work related to this?} Nevertheless, we leave these considerations for later works.
 
@@
 
-To derive the first and the second law we need to give a definition for work and heat.
-We do so by defining two special thermodynamic systems.
+A natural question, based on the previous definition, is: to what extend thermodynamic processes can change the energy and the entropy of each system?
+Intuitively, there are two ``orthogonal'' directions for a change of energy: constant entropy and maximal entropy change.
+To study these components, it is useful to consider the following special thermodynamic systems.
+
 We say a thermodynamic system $R$ is a \textbf{thermal reservoir} if its equation of state can be sufficiently approximated by $S = \beta U = \frac{U}{k_B T}$.
@@
 
-We say $M$ is a mechanical system if the state entropy is a constant for all states.
+We say $M$ is a \textbf{mechanical system} if the state entropy is a constant for all states.
 For a thermodynamic system that is also mechanical, we therefore have  $S(U, x^i) = k_M$ for all its states.
@@
 The energy of the mechanical system is effectively \emph{stored} in one of the other state variables and can be later retrieved.
-We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U$ as the energy gained by the system during a thermodynamic process.
+We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U$ as the energy gained by the system during a thermodynamic process.\footnote{Note that $\frac{\partial S_M}{\partial U} = 0 = \beta_M = \frac{1}{k_B T_M}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite.
+While this ``temperature'' has little to do with thermometers, the fact that we can always transform all work into heat but not vice-versa is just a specific case of being able to move energy from higher to lower temperature.}
 
-These two types of systems allow us to study the two ``orthogonal'' directions a change of energy can have: an energy change in the direction of constant entropy and one in the direction of maximal energy change.
-These definitions makes this manifestly clear.
-Also note that $\frac{\partial S_M}{\partial U} = 0 = \beta_M = \frac{1}{k_B T_M}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite.
-While this ``temperature'' has little to do with thermometers, the fact that we can always transform all work into heat but not vice-versa is just a specific case of being able to move energy from higher to lower temperature.
+We can study how changes of energy and entropy are related to each other.
+Consider a system composed of a generic system $A$, a thermal reservoir $R$ and a mechanical system $M$.
+In this setting, work represents the energy exchanged at constant entropy while heat represents energy exchanged associated with entropy change.
 
-We now consider a system composed of a generic system $A$, a thermal reservoir $R$ and a mechanical system $M$.
 During a thermodynamic process, because the total energy of the system must be conserved, we have:
@@
 \end{equation}
-This recovers the first law.
+This recovers the first law of thermodynamics.
 
@@
 
-As state entropy counts the possible evolutions of the system at equilibrium, the lowest possible count $\mu$ is one, which is achieved if the equilibrium can be reached only by one starting condition.
+Recall that state entropy counts the logarithm of the possible evolutions of the system at equilibrium.
+The lowest possible count $\mu$ of evolution is one, which is achieved if the equilibrium can be reached only by one starting condition.
 In this case, $S=\log(\mu)=0$.
@@
 Consequently, the minimum for entropy is reached either at the lowest or highest energy state.
-Now we show that all thermodynamic systems must have either a lowest energy state or a highest energy state.
+Now we show that either all thermodynamic systems must have a lowest energy state or they must have a highest energy state.
 Suppose, in fact, we had two system, $A$ has no maximum energy while $B$ has no minimum energy.
@@
 
-\section{Discussion}
+\section{Connections to mechanics}
 
-We have shown that thermodynamics can be conceptually rederived by adding few other assumptions/definitions from a general characterization of states, processes and equilibria.
-Can this approach also offer new insights on classical mechanics, quantum mechanics and statistical mechanics that may lead to a more unified understanding these somewhat disconnected disciplines?
+While recovering thermodynamics from straight forward physical definitions has its own merits, we believe the approach can be used to offer gave additional insights new insights on classical mechanics, quantum mechanics and statistical mechanics.
+So we want to conclude with a few remarks on the subject that give a feel for how we intend to extend the work in the fully developed framework.
 
-The notion of equilibria would be a place to start.
-We defined, as it is customary, the state space of a thermodynamic system as consisting only of states of equilibria.
-We also defined a mechanical system as one for which $S$ is the same for all states.
-But we said the state entropy is uniquely defined only for state of equilibria.
-This implicitly assumed that all states of a mechanical systems are equilibria.
-Is this always true?
-And in what sense is this true?
-
-Let us first consider the classical phase space of a particle.
-Qualitatively, we can imagine saying that a particle at rest is in equilibrium.
-A particle in uniform motion, though, does not seem in equilibrium: the distance is constantly increasing.
-Yet, there exists a frame in which it is at rest.
-So, in this a sense, all free particles are at equilibrium.
-Going one step further, in the presence of gravity, a particle at rest in one frame will not be in uniform motion in another.
-So it would seem that a particle under gravitational or apparent forces is at equilibrium.
-What is going on?
-The issue here is that, as we noted before, the notion of equilibrium as ``nothing is changing'', which is the notion coming from dynamical systems, is not exactly aligned with the notion of equilibria we need for thermodynamics.\footnote{XXX: cite papers that note the difficulty of definiting temperature relativistically} So we need to answer two questions: what would be a more suitable definition?
-And why would that definition be tied to the notion of state?
-
-In another work, we more thoroughly investigate characterizing equilibria in terms of system decoupling, which we briefly present here.
-By system decoupling we mean a process that makes the system independent from other systems, its evolution does not constrain and is not constrained by the evolution of other systems.
-As we saw before, independence was already critical for defining state entropy.
-But why should it be connected to states?
-We typically think of a state as the description of one system and, implicitly, only that system.
-If the system is not independent, if there are correlations, the description of that system tells us something about the other as well.
-So if we want a state to give the description of one system alone, we implicitly describing it in the case where the system is independent.
-System independence is not just a requirement for the process entropy to factorize: it's an implicit requirement to properly talk about a system and its state.
-But why would independence be tied to a notion of equilibria?
-The idea is that when we group together quantities into a system, this is not done arbitrarily.
-For example, we do not group together the position of one ball with the velocity of another.
-We group quantities that ``go together'', that have to be studied together.
-More precisely, the ability to define a system impinges upon the ability to prepare it and study it independently from the rest.
-If we are not able to then what we have is not really a system, just a set of random quantities.
-So, at least in line of principle, we must have a way to decouple the system from other systems.
-We must have a process that makes the system independent.
-Clearly, if the system is already independent, the process should do nothing.
-If states are descriptions of an independent system, they must be equilibria of these decoupling processes.
-In this sense, we can say that all states, thermodynamic equilibria or not, are equilibria.
-Thermodynamic equilibria extends this notion of independence at different level of scale.
-
-Still, one may find strange associating a state entropy to mechanical system.
-And why would that be constant on all states?
-We should note that the idea that all states have the same entropy is already embedded in statistical mechanics.
+We have argued the states of a system are implicitly independence equilibria and have an associated state entropy.
+One does not typically speak in these terms for classical and quantum states.
+Is this warranted and useful?
+Do all mechanical system match our requirement to have constant entropy?
+If one looks closely, these facts are implicit assumed.
 For quantum mechanics, in fact, the Von Neumann entropy of every pure state is zero.
-In classical mechanics, a uniform measure is used to calculate the accessible states; additionally the Gibbs entropy is invariant under translations in phase space.
+In classical mechanics, a uniform measure is used to count the accessible states; additionally the Gibbs entropy is invariant under translations in phase space.
 It is also embedded directly in the corresponding mechanics.
-In both theories, given to states, we can always (at least mathematically) find a deterministic and reversible evolution (i.e.
+In both theories, given two states, we can always (at least mathematically) find a deterministic and reversible evolution (i.e.
 a Hamiltonian evolution or a unitary evolution) that connects the two.
@@
 Entropy as the count of microstates is easily recovered.
+Yet, it is recovered under specific conditions.
 
-Note that this can happen in different scenarios.
+These conditions can be met in different scenarios.
 We could have an independent mechanical system under a deterministic and reversible process.
@@
 This example also show how critical is the notion of independence to define state entropy.
-The evolution of a damped harmonic oscillator is not independent from the evolution of the systems that are absorbing the dissipated energy.
+The evolution of a damped harmonic oscillator is not independent from the evolution of the systems that are absorbing the dissipated energy, which are in fact correlated: the higher the energy of the oscillator, the higher the energy lost.
 The same description at different times (e.g.
@@
 
-\section{Conclusion}
+\section{Conclusion and further work}
+
+In this paper we gave a conceptual overview for a new approach to recover thermodynamics based on the simple idea of counting evolutions, the possible ways a system can evolve under a process.
+We saw that the logarithm of evolution count under a deterministic process has all the properties we associate with entropy and can serve as a foundations for thermodynamics but also as a common foundations for the other physical theories.
+
+Our aim is to fully develop a formal framework based on these ideas, which will give us a general theory for states and processes.
+Conceptually, we start with all that can be observed at all moments of time; organize them into systems/degrees of freedom by grouping those that ``go together'' in the sense that they can be studied, at least under some circumstances, independently from the rest; define state spaces as ``templates'' of those descriptions that can be instantiated over time.
+This will force us to spell out the hidden assumptions one always have when defining systems, states and processes.
+The goal is to then recover the different physical theories with additional assumptions on the system or on the processes being studied.
+
+Mathematically, we will build on our work on the logic of experimental verifiability, that already gives us the basic building blocks (i.e.
+topologies, $\sigma$-algebras).
+This need to be extended to allow us to count possibilities in general and evolutions in particular.
+Technically, we need a unified tool that generalizes notions of measure theory, geometry, probability and information theory as these cover specific aspect that do not work in the general case.
+
+It is important to us that these additional efforts, while making the mathematics more precise, will not obscure the physical motivation and intuitive understanding of the framework.
+Mathematical consistency is a tool We believe there is a way
+
+These additional efforts will not change the nature of the insights presented.
 

### 2021-01-05 Articles/2019-ShannonEntropy/Submissions/EJP/response.tex
Reviewed before sending to Julian

@@
 
-	Thanks editor.
+We thank the editor for the opportunity to revise and resubmit this article based on the feedback of the referees. We hope that the improved article is now suitable for publication in EJP.
+
+Sincerely,
+
+Christine A. Aidala, Gabriele Carcassi, Julian Barbour
 
@@
 
-We find the attacks on the reference completely out of place, inappropriate and distasteful. Ad hominem attacks have no place in the scientific discourse.
+We find the attacks on the reference completely out of place, inappropriate and distasteful.
+Ad hominem attacks have no place in the scientific discourse.
+However, we have removed the ``offending'' reference and replaced it with another one, now from the physics education literature.
 
@@
 ``Entropy'' is used for the physical concepts in statistical mechanics.
-That helps the reader understand that some properties of that concept are more general than physics, which is the proper setting to understand the concept.
+That helps the reader understand that many properties of the Shannon entropy are more general than their specific use in physics.
+And the more general setting is the proper one to understand the concept.
 Given that we make a distinction between the two words, ``variability'' and ``entropy'' are \textbf{not} synonyms in our work.
@@
 Sheldon Goldstein et al.
-In many modern textbooks of statistical mechanics, the Shannon entropy is not the starting point, rather an approximation that is valid only at equilibrium.
+Moreover, in many modern textbooks of statistical mechanics, the Shannon entropy is not the starting point, rather an approximation that is valid only at equilibrium.
 See, for example, ``Statistical mechanics in a nutshell'' by Luca Peliti, where the starting point is state counting.
@@
 Doing a comparative analysis of this kind is beyond the scope of the paper, which is simply to introduce in a clear way to newcomers what the expression $- \sum p \, \log p$ represents.
-See, for example, "Compendium of the foundations of classical statistical physics" by Jos Uffink.
+See, for example, "Compendium of the foundations of classical statistical physics" by Jos Uffink for an example of such comprehensive review.
 
@@
 Again, we are in no way disproving what Jaynes says.
-What the referee (but he is not the only one) is engaging in what philosophers define as ``verbal dispute'', where one disagrees on the meaning of the words rather than on an actual fact of the matter.
+The referee (but he is not the only one) is engaging in what philosophers define as ``verbal dispute'', where one disagrees on the meaning of the words rather than on an actual fact of the matter.
 The problem is what one means by ``subjective''.
@@
 
-We have expanded the section to clarify that and also cited the 1992 which does not change anything. [TODO include]
+We have expanded the section to clarify that and also cited the 1992 paper, which however does not affect our point:
+
+\emph{We note that some authors choose to interpret the probability distribution not as coming from repeated independent trials, but as the knowledge one has about the system.
+This would make the entropy a subjective notion: each observer would have a different credence distribution, regardless of whether it fits the data, and therefore a different Shannon entropy.
+A physical quantity, however, must be the same for everybody, and the associated fluctuations we experimentally observe are indeed objective.
+So what is going on?}
+
+\emph{We believe the confusion comes from the, correct, realization that the same system, under different conditions, will be described with a different set of thermodynamic variables.
+The choice of the system, in statistical mechanics and thermodynamics, is enough to determine the state space for the microstates, but not enough to determine the set of macrostates that correspond to equilibria.
+We have to specify the process and the constraints that that process puts on the system.
+Under a different choice of process and constraints the microstates will fluctuate in different ways since we have changed the dynamics of the system.
+This is what we stated at the beginning: a distribution, and therefore its variability, is always contingent upon some arbitrary choices.
+Since the Gibbs entropy is the variability of a microstate within the distribution identified by the macrostate, it is not a property of the single microstate, it is not a property of the system, but it is a property of the system within that specific process, of the macrostate.
+Some authors refer to this fact by saying that entropy is not objective.}
+
+\emph{The issue is that subjective, in the context of probability theory, refers to Bayesian probability, which really means subjective: what one believes to be true regardless of what experimental evidence there is.
+This is not at all the same concept as the one outlined before.
+The system plus the process (which identifies the set of constraints) determines the ensemble and therefore the entropy, regardless of whether an agent knows what process was used.
+Even if one wants to give a Bayesian account, then, one has to give it in terms of an agent that has full and exact knowledge of the system and the process.
+But this is hardly subjective.
+In our view, this is a case of unfortunate word choice.
+It is much better to simply state that the entropy is a property of the ensemble, of the distribution, and not of the system.
+And if one understands that the Gibbs entropy is the variability of the microstate within the ensemble, then it is clear that it cannot be a property of the microstate itself, but it is a property of the ensemble.}
 
@@
 
-Given the complete mis-characterization of the prose and the erroneous claim that somehow the choice of unit/granularity makes the Shannon entropy always positive, we find it impossible to understand whether there is any meaningful item to address.
+Given the complete mischaracterization of the prose and the erroneous claim that somehow the choice of unit/granularity makes the Shannon entropy always positive, we find it impossible to understand whether there is any meaningful item to address.
 One possible answer is that the referreee does not realize that choosing a unit, in the context of information entropy, is equivalent to choosing the resolution since it changes the zero for entropy...
@@
 They are called \textbf{defining constants} because they define the unit system, which is a convention.
-Before engaging in comments that imply we are incompetent, the referee may want to read the relevant literature.
+Before engaging in comments that imply we are incompetent, the referee may want to read the relevant literature, which are now included in the paper.
 
-We added the following footnotes, merely as a pointer for the curious reader. [TODO]
+We added the following footnotes, merely as a pointer for the curious reader:
+
+\emph{In the latest revision of the International Metric System of Units SI, the Boltzmann constant is one of the defining constants.
+This means its value is assigned, rather then measured, in a way that properly defines Kelvin, the unit of temperature.
+The universal constants are essentially used as fixed starting points to define our measurement scales.}
+
+\emph{As [...] notes:
+\begin{quote}
+	Temperature characterizes the average thermal energy of particles in a certain ensemble of particles in the state of thermodynamic equilibrium.
+Therefore, in principle, one could express this quantity in terms of the unit of energy; this is actually practiced in many fields of physics, especially in statistical physics, where it is quite common to use $\Theta = k_B T$ or $\beta = 1 / k_B T$  as a measure of thermal energy, instead of temperature $T$ measured in Kelvins.
+\end{quote}}
 
@@
 
-This referee report, while abundant in ad hominen, mischaracterizations and demeaning language, was outstandingly lacking in constructive criticism.
+This referee report, while abundant in ad hominem, mischaracterizations and demeaning language, was outstandingly lacking in constructive criticism.
 This is peer reviewing at its worst.
-The most likely scenario is that, after a very cursory look, he framed the manuscript in some diatribe the referee has with some other colleagues, stopped reading carefully and engaged in a tirade, assigning to us views that are not our own.
-This means that most of the comments weren't even directly relevant to our manuscript: even after sifting the report carefully, we found few actionable items.
+The most likely scenario is that, after a very cursory look, he framed the manuscript in terms of some dispute the referee has with some other colleagues, stopped reading carefully and engaged in a tirade, assigning to us views that are not our own.
+This means that most of the comments weren't even directly relevant to our manuscript: even after sifting the report carefully, we found no major actionable items.
 
@@
 
-Most of the literature we found in Physics Education focuses on the Gibbs, Boltzmann and Thermodynamic entropies.
+Most of the literature we found in Physics Education focuses on the Gibbs, Boltzmann and thermodynamic entropies.
 We added references for those in the first paragraph of the introduction, along with a remark stating that the Shannon entropy itself has received less attention.
@@
 \end{quote}
+
+Thank you, this has been fixed.
 
@@
 
---------------------
-
-Closing.
-
-Sincerely,
-
-The Authors
-
 \end{document}

### 2021-01-05 Articles/2019-ShannonEntropy/Submissions/EJP/EJP-ShannonEntropyIsVariability.tex
Adding link to youtube videos
Reviewed before sending to Julian

@@
 \usepackage{graphicx} % needed for figures
+\usepackage{hyperref}
 
@@
 
-The goal of this work is to give a comprehensive introduction to the concept of Shannon entropy, the expression $-\sum p_i \log p_i$, which is well established in such disparate fields as computer science, communication theory, physics\cite{Jaynes1957-1, Grandy} ecology and economics.
-The general concept of entropy in physics, despite being treated extensively in the physics education literature (e.g.
-[TODO cite physed]), is still surrounded by a great deal of confusion [TODO cite evidence of confusion]\cite{Swendsen, KishFerry}.
-Partly because, at a fundamental level, there is still disagreement on what is the ``correct'' notion of entropy, particularly out of equilibrium.
-[TODO cite works of out of equilibrium research] Partly because the same word ``entropy'' is used to refer to concepts (e.g.
-topological entropy, thermodynamic entropy, graph entropy, R'{e}nyi entropy, information entropy) that have different definition and various degree of overlap.
+The goal of this work is to give a comprehensive introduction to the concept of Shannon entropy, the expression $-\sum p_i \log p_i$, which is well established in such disparate fields as computer science, communication theory, physics \cite{Jaynes1957-1, Grandy} ecology and economics.
+The concept of entropy in physics, despite being treated extensively in the physics education literature (e.g.
+\cite{LeffPartI, LeffPartII, LeffPartIII, LeffPartIV, LeffPartV, carson2002undergraduate, baierlein1994entropy, geller2014entropy}), is still surrounded by a great deal of confusion \cite{Swendsen, Styer}.
+Partly because, at a fundamental level, there is still disagreement on what is the ``correct'' notion of entropy, particularly out of equilibrium \cite{jarzynski2017stochastic, goldstein2019gibbs, maes2003time}.
+Partly because the same word ``entropy'' is used to refer to concepts (e.g.
+topological entropy, thermodynamic entropy, graph entropy, R\'{e}nyi entropy, information entropy) that have different definitions and various degrees of overlap.
 We will focus on the notion of Shannon entropy, by which we mean the use of $- \sum_i p_i \log p_i$ in the broadest characterization, which has received less attention in physics education.
@@
 The contexts are different and relating them goes beyond the scope of this work.
-}
+} The same material is also presented in a series of videos \cite{videos}.
 
@@
 We note that some authors choose to interpret the probability distribution not as coming from repeated independent trials, but as the knowledge one has about the system.
-This would make the entropy a subjective notion: each observer would have a difference credence distribution, regardless of whether it fits the data, and therefore a different Shannon entropy.
-A physical quantity, however, must be the same for everybody and the associated fluctuations we experimentally observe are indeed objective.
+This would make the entropy a subjective notion: each observer would have a different credence distribution, regardless of whether it fits the data, and therefore a different Shannon entropy.
+A physical quantity, however, must be the same for everybody, and the associated fluctuations we experimentally observe are indeed objective.
 So what is going on?
@@
 
-The issue is that subjective, in the context of probability theory, refers to Bayesian probability, which really means subjective: what ones believes to be true regardless of what experimental evidence there is.
+The issue is that subjective, in the context of probability theory, refers to Bayesian probability, which really means subjective: what one believes to be true regardless of what experimental evidence there is.
 This is not at all the same concept as the one outlined before.\footnote{Note how Jaynes\cite{Jaynes2} always put ``subjective'' in double quotes.} The system plus the process (which identifies the set of constraints) determines the ensemble and therefore the entropy, regardless of whether an agent knows what process was used.
@@
 It is much better to simply state that the entropy is a property of the ensemble, of the distribution, and not of the system.
-And if understand that the Gibbs entropy is the variability of the microstate within the ensemble, then it is clear that it cannot be a property of the microstate itself, but it is a property of the ensemble.
+And if one understands that the Gibbs entropy is the variability of the microstate within the ensemble, then it is clear that it cannot be a property of the microstate itself, but it is a property of the ensemble.
 
@@
 
+\section{Acknowledgments}
+We would like to thank Juniar Lucien for insights into the physics education literature.
+This work is in connection to Assumptions of Physics, a larger project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
+
 \bibliographystyle{unsrt}  

### 2021-01-05 Articles/2019-ShannonEntropy/Submissions/EJP/EJP-ShannonEntropyIsVariability.tex
Fix bibliography formatting and add three EJP physics education references

@@
 The concept of entropy in physics, despite being treated extensively in the physics education literature (e.g.
-\cite{LeffPartI, LeffPartII, LeffPartIII, LeffPartIV, LeffPartV, carson2002undergraduate, baierlein1994entropy, geller2014entropy}), is still surrounded by a great deal of confusion \cite{Swendsen, Styer}.
+\cite{baierlein1994entropy, carson2002undergraduate, Santillan_2008, LeffPartI, LeffPartII, LeffPartIII, LeffPartIV, LeffPartV, Benguigui_2013, geller2014entropy, Tiwary_2020}), is still surrounded by a great deal of confusion \cite{Swendsen, Styer}.
 Partly because, at a fundamental level, there is still disagreement on what is the ``correct'' notion of entropy, particularly out of equilibrium \cite{jarzynski2017stochastic, goldstein2019gibbs, maes2003time}.

### 2021-01-06 Overviews/Classical mechanics/ClassicalMechanicsOverview.tex
Minor improvement in classical overview

@@
 This means the volume form is proportional to $\omega^n$.
-This cannot be degenerate (i.e.~it must be nonzero for each infinitesimal volume), or some regions of the particle state space would not contain states, which is a contradiction.
+This cannot be degenerate (i.e.~it must be nonzero for each infinitesimal volume) since all regions of phase space must, by definition, contain some states.
 Therefore $\omega$ itself cannot be degenerate.

### 2021-01-07 Articles/2019-Entropy/Source/thermodynamics.tex
Adding another figure in thermodynamics paper

@@
 This will not happen, however, if the systems are independent.
-That is, for each possible evolution $x(t)$ of the first system and $y(t)$ of the second system, there will be a possible evolution $\{x(t), y(t)\}$ on the composite.
+That is, as shown in \ref{fig_independence}, for each possible evolution $x(t)$ of the first system and $y(t)$ of the second system, there will be a possible evolution $\{x(t), y(t)\}$ on the composite.
 In other words,
@@
 the measure factorizes. We say two system are \textbf{independent} at given time $t$ if the evolutions per state factorizes: $\mu_{X \times Y}(x(t), y(t)) = \mu_X (x(t)) \mu_Y (y(t))$.
+
+\begin{figure}[h]
+	\includegraphics[width=\columnwidth]{images/Slide7.png}
+	\caption{Independent system.
+If two systems are independent, the evolution of one does not restrict the evolution of the other: all pairs are possible.
+The number of evolutions factorizes.}\label{fig_independence}
+\end{figure}
 

### 2021-01-11 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
Extensive edits to Section 4: rewriting intro, rewriting section 4.2 (incorporating gauge redundancy argument from section 5.2); edits to 4.1 and 5.2.

@@
 For instance, we can think of the canonical position and momentum observables as ``epistemically fundamental" because they suffice to completely specify a classical state, fixing its values for all other observables.
-\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{density},discusses another sense in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied.
+\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{density}, discusses another sense in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied.
 Compared with North's \parencites*[]{North} account of structure, our definition has two distinct advantages.
@@
 
+\begin{comment}
 * ``Fundamentality" because it makes the physics more prominent.
@@
 
-\subsection{Density and Riemmanian geometry}
-
-It is helpful to take a quick look at how densities work in general so that we can better understand what is so special about them in the context in Hamiltonian mechanics.
-By density we mean the intuitive physical notion of a quantity divided by the region it occupies, like mass per volume or charge per area.
+\end{comment}
+
+We have now shown that Hamiltonian mechanics requires less structure than Lagrangian mechanics because it requires fewer physical conditions.
+While this provides one reason for privileging Hamiltonian mechanics, it is not the only one.
+A special relationship exists The special relationship between Hamiltonian mechanics and invariant densities provides a further---no less important---reason for privileging Hamiltonian mechanics.
+In this section, we argue that viewing physics through the lens of invariant densities has much to commend it.
+In general, it provides a better understanding of classical systems.
+
+In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
+If the scalar value of the density at one point is larger than another, this entails that the density is higher at that point of phase space.
+Physically, this means that there is more physical stuff at this point.
+Likewise, if the scalar value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
+For brevity, will refer to this as the property of \textit{point-wise comparison}.
+Point-wise comparison requires a coordinate-invariant density.
+Defining an invariant density requires, in turn, that we define state space over a symplectic manifold, as the derivation in Section~\ref{deterministic} shows.
+In other words, to define a density with the property of point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
+
+Section~\ref{dependence} shows why point-wise comparison does not hold in general: the density is a coordinate-dependent quantity, and thus it is not generically coordinate-invariant.
+Consequently, within the Lagrangian formalism, point-wise comparison holds only within special coordinate systems.
+This gives a further reason to privilege Hamiltonian mechanics.
+Section~\ref{fundamental} argues for the fundamental physical significance of the phase space density.
+Since the density is coordinate-dependent, this shows that coordinate-independence is not a necessary condition for fundamental physical significance.
+Our position undercuts metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms [get citations: Field 1980; Chen papers; North].
+Additionally, we explain how the conjugate momentum $p$ is itself a gauge-dependent quantity, bolstering our claim that the density is fundamental, not the state variables.
+Hence, the real physical significance of canonical coordinates is that they allow us to define an invariant phase space density $\rho $.
+
+\subsection{Coordinate-dependence and coordinate-variance}
+\label{dependence}
+
+%It is helpful to take a quick look at how densities work in general so that we can better understand what is so special about them in the context in Hamiltonian mechanics.
+
+To understand what makes densities special in the context of Hamiltonian mechanics, it is helpful to consider densities in general.
+By a `density' we mean the intuitive physical notion of a quantity divided by the region it occupies, like mass per volume or charge per area.
 Over a continuum, this is an integrable function $\rho(x)$.
-Mathematically, the concept is captured by the Radon–Nikodym derivative: we have a region $U$, we have two measures $m$ and $V$ and the derivative satistifies $m(U) = \int_U \rho dV$.
-This makes it clear that the units of $\rho$ depends on the units of both $m$ and $V$.
-More precisely, if $\rho(x, y)$ is a function of multiple variables, we will want $\int_U \rho(x, y) dx dy$ so that the units of $\rho$ are consistent with the units of the arguments.
-For example, in polar coordinates $\rho(r, \theta)$ will return mass over meter-radians if we measure $r$ in meters and $\theta$ in radians.
+Mathematically, the concept is captured by the Radon–Nikodym derivative: we have (i) a region $U$, (ii) two measures $m$ and $V$, and (iii) the derivative satistifies $m(U) = \int_U \rho dV$.
+This relationship shows that the units of $\rho$ depend on the units of both $m$ and $V$.
+More precisely, if $\rho(x, y)$ is a function of multiple variables, dimensional analysis requires that the units of $\rho$ are consistent with the units of the arguments in the expression $\int_U \rho(x, y) dx dy$.
+In polar coordinates $(r, \theta)$, if we measure $r$ in meters and $\theta$ in radians, then $\rho(r, \theta)$ must return mass over meter-radians .
 
@@
 Worse than that, it is not even coordinate invariant: the density changes with the Jacobian.
-For example, changing from polar to Cartesian coordinates we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
-Since the Jacobian is a function of the point, this makes the value of the density, in general, not directly comparable at different points.
-If we have a uniform density in $x$ and $y$, we will have an $r$-dependence in $r$ and $\theta$.
+For example, changing from polar to Cartesian coordinates, we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
+Since the Jacobian is a function of the point, the value of $|J|$ can change from point to point.
+Consequently, the density is not guaranteed to be directly comparable at different points: point-wise comparison need not hold in an arbitrary coordinate system.
+If we have a uniform, constant density in $x$ and $y$, we will have an $r$-dependence in $r$ and $\theta$.
 This dependence is introduced precisely because the area identified by $dr$ and $d\theta$ is $dr \,r d\theta$, which is a function of $r$.
+This makes the value of the density, in general, not directly comparable at different points.
+Hence, to determine whether there is more stuff at one location in state space than another, we need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
+we can't just compare the values of $\rho$ in these coordinates and see which is larger).
+%The Hamiltonian formalism circumvents this further epistemic requirement by defining the density using quantities that vary covariantly.
+Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
+%when we define the density originally, it is coordinate-free (rho as a function of the state space of the particles, S).
+But when we go to express it numerically, then it becomes coordinate-dependent.
+(I.e.
+when we express it over a manifold, then it needs to be a density for integration, and hence it becomes coordinate-dependent).
+[see 57.3 minutes of 12-21 recording for discussion]
 
@@
 Considering the total mass in the region $U$, we have $m(U) = \int \nu$ so essentially $\nu = dm$ which is the infinitesimal mass contribution we are integrating.
-But this means the value returned by $\nu$ depends not only on how must stuff there is at a given point, but also on the size of the volume identified by the $n$ vectors given to the form.
-[Not immediately clear to me what the upshot is of this last point; e.g.
-how does this relate to $\nu$ being coordinate-free?]
+But this means that the value returned by $\nu$ depends not only on how much stuff there is at a given point, but also on the size of the volume identified by the $n$ vectors given to the form.
+In other words, $\nu$ has dimensions of only mass and is defined as a function of both a point and a volume element.
+In contrast, the density $\rho$ is defined at each point and has dimensions of mass over volume.
+This shows that the coordinate-independent function $\nu$ cannot play the role of the coordinate-dependent density $\rho$.
 
@@
 a_n}$ is the Levi-Civita symbol.
-We can then express any differential density $\nu = \hat{\rho} dV$ as the product of the volume density times a scalar function $\hat{\rho}$.
-At first glance, this expression misleadingly suggests that the density $\nu$ is coordinate-independent, but this is not the case.
-[double check that this claim applies to $\nu$ rather than $\hat{\rho} $.
-Is this in tension with saying earlier that $\nu$ is coordinate-free?] The quantity $|g_{ab}|$ depends on the choice of coordinates and is, in general, a function of the point.
+We can then express any differential density $\nu = \hat{\rho} dV$ as the product of the volume density times a scalar function $\hat{\rho} \equiv \rho \ \epsilon \sqrt{|g_{ab}|}$.
+At first glance, this expression misleadingly suggests that the density $\hat{\rho}$ is coordinate-independent: it is a scalar quantity with the correct dimensions of mass over volume.
+Nevertheless, $\hat{\rho}$ remains coordinate-dependent because it is a function of $|g_{ab}|$, which depends on the choice of coordinates and is, in general, a function of the point.
 However, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
-The determinant $|g_{ab}|$ is the Jacobian of the transformation to those coordinates [the rationale/point of this sentence was not clear to me].
-For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$; in polar coordinates $|g_{ab}| = r^2$ and the volume form $dV = r dr \wedge d\theta$ [above we said that the Jacobian was $r$ rather than $r^2$].
-This means that $\hat{\rho}$ is really expressing the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise.
-We can also see this if we look at the units of $\hat{\rho}$, which are in terms of inverse volume as expressed in the privileged system.\footnote{This is similar to what happens with the concept of proper time: it is defined in the comoving frame, so it is coordinate dependent.
-Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate invariant.}
-
-We can now look at what happens for distributions over position and velocity.
+Since the square root of the determinant $|g_{ab}|$ is the Jacobian of the transformation to those coordinates, $|g_{ab}|$ will be constant only when this Jacobian equals one.
+For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$; in polar coordinates $|g_{ab}| = r^2$ and hence the volume form $dV = r dr \wedge d\theta$.
+This means that $\hat{\rho}$ at best expresses the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise.\footnote{This is similar to what happens with the concept of proper time: it is defined in the comoving frame, so it is coordinate dependent.
+Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate invariant.} Again, the upshot here is that we cannot avoid using a coordinate-dependent density.
+Outside the context of Hamiltonian mechanics, the best we can do is obtain the property of point-wise comparison in a special coordinate system.
+%We can also see this if we look at the units of $\hat{\rho}$, which are in terms of inverse volume as expressed in the privileged system.
+From 1-4 meeting: I thought that $\hat{\rho}$ has units of mass over volume.
+
+We can now look at what happens for distributions over position and velocity, i.e.
+on the tangent bundle in Lagrangian mechanics.
 Note that the velocities transform in the same way (i.e.
 contravariantly) as differentials in position.
-Therefore $\rho(x^i, v^j) = \sqrt{|g_{ab}|} \sqrt{|g_{ab}|} \rho(\hat{x}^i, \hat{v}^j) = |g_{ab}| \rho(\hat{x}^i, \hat{v}^j)$ $dx^a$.
-The determinant of the metric tensor is therefore directly the volume of the tangent bundle [I didn't follow this point].
+Consequently, the density $\rho(q,\dot{q})$ also varies contravariantly.
+These transformation properties preclude the density---in the Lagrangian formalism---from being coordinate-invariant in general.
+In detail, $\rho(x^i, v^j) = \sqrt{|g_{ab}|} \sqrt{|g_{ab}|} \rho(\hat{x}^i, \hat{v}^j) = |g_{ab}| \rho(\hat{x}^i, \hat{v}^j)$.
+The determinant of the metric tensor is therefore directly the volume of the tangent bundle.
 The relationship to the density as expressed in phase space is therefore the following:
@@
 
----
-
-Section on Ham. Not everything physical has to be coordinate independent or invariant.
-
-* In phase space, we do have coordinate invariance and point-wise comparison.
-Momentum changes covariantly.
-It is the density that is truly fundamental, not the coordinates in general.
-Though, the q's are priviledged from epistemic access.
-The p's are not, in fact we can define gauge transformation as things that keep the q's the same but change the p.
-
-*[Could save the following for discussion of gauge redundancy in section 5.2] Not all ``state variables" are physical.
-Hamiltonian mechanics does not really care what the q's and p's are.
-It is the motion assumption that tells us that q's are position, simply because they define the unit system.
-P is a guage quantity (we care about the trajectory, the velocity).
-In this view, p is simply a ``label'' we can use to express the density over states.
-Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves.
-$->$ Invariant density is really the important physical element/object in Hamiltonian mechanics.
-
-* Units are physically significant.
-In fact, the ability to express the densities in units is what gives us the symplectic structure.
-It's a physical requirement that we must be able to express the density in some unit.
-
-* Invariant densities: physical significance connected to coordinate invariance. [incorporated]
-
-* Density comparison is independent of coordinates. We can compare densities point-wise without worry. [incorporated]
-
-* Makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
-Areas/volumes are measures of number of states.
-They are coordinate invariant.
-Also the Shannon/Gibbs entropy are coordinate invariant [but not coordinate independent?] (see physicality of quantities that are coordinate invariant)
-
-* Close the section with a more general discussion on physicality/fundamentality and all those other physical goodness.
-
-We have now substantiated the claim of Section~\ref{privileging} that Hamiltonian mechanics requires less structure than Lagrangian mechanics, in virtue of requiring fewer physical conditions.
-While this provides one reason for privileging Hamiltonian mechanics, it is not the only one.
-A further---possibly even more important---reason stems from the relationship between Hamiltonian mechanics and invariant densities.
-As we argue in this section, viewing physics through the lens of invariant densities has much to commend it.
-In general, it provides a better understanding of classical systems.
-Specifically, we will see that densities are of fundamental physical significance due to their coordinate-invariance.
-Hamiltonian mechanics---unlike Lagrangian mechanics---allows us to compare these densities point-wise in any coordinate system.
-Although the density is coordinate-invariant, it is not coordinate-independent (it cannot be given a coordinate-free definition).
-This is because expressing the density requires a choice of units, a fact that is closely connected with symplectic structure.
-Hence, we will argue that some coordinate-dependent representational choices are nevertheless physically significant.
-This view is at odds with metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms [get citations: Field 1980; Chen papers; North].
-% Section~\ref{action} uses the invariant density to provide a geometrical interpretation of the action principle.
+Overall, these considerations show that in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
+In this context, we need to know which coordinate system we are in before making these comparisons.
+The Hamiltonian formalism circumvents this further epistemic requirement by defining the density using quantities that vary covariantly.
+Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
+% Josh: why is the determinant of the metric tensor equal to the volume of the tangent bundle in this case? [Discussed on 1-4 meeting, 11:56 a.m.]. Perhaps study discussion of this in the long paper.
+
 % Remember: the density is coordinate-invariant but not coordinate-independent; it is actually coordinate-dependent (i.e. it does not admit of something like a coordinate-free definition)
-% Also remember: there are quantities that are both coordinate-invariant and coordinate -dependent that are still physically significant, e.g. choices of q's and p's (?)
+% Also remember: there are quantities that are both coordinate-variant and coordinate-dependent that are still physically significant, e.g. choices of q's and p's (?)
 % Coordinate-invariant in general means that definition of the quantity involves referencing coordinates, but if you change the coordinates, then the value of the quantity remains the same.
@@
 
-In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
-If the scalar value of the density at one point is larger than another, this entails that the density is higher at that point of phase space.
-Physically, this means that there is more physical stuff at this point.
-Likewise, if the scalar value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
-For brevity, will refer to this as the property of \textit{point-wise comparison}.
-Since in general the density is coordinate-dependent, this property of point-wise comparison does not always hold.
-For instance, it can fail as we change from Cartesian to polar coordinates [figure out where to insert or reference the illustration of this shown above].
-For point-wise comparison to hold, we require a coordinate-invariant density.
-As the derivation in Section~\ref{deterministic} shows, we must define our state space using a symplectic manifold in order to define an invariant density.
-In other words, to define a density with the property of point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
-As we will discuss further in Section~\ref{gauge}, the conjugate momentum $p$ is itself a gauge-dependent quantity.
-Hence, the real physical significance of canonical coordinates is that they allow us to define an invariant phase space density $\rho $.
-[Move this section 5.2 stuff up into this section; and figure out how to reorganize this paragraph]
-
-It is instructive to contrast this situation with Lagrangian mechanics.
-In this case, the density depends on the position and velocity state variables $(q,\dot{q}) $.
-Crucially, these state variables transform contravariantly, rather than covariantly.
-Consequently, the density $\rho(q,\dot{q})$ also varies contravariantly.
-These transformation properties preclude the density---in the Lagrangian formalism---from being coordinate-invariant in general.
-Of course, there exist special coordinate systems that allow for faithful point-wise comparison of the density.
-For instance, Cartesian coordinates---for a system in an inertial frame of reference---support this formal representational property.
-Yet, we still need to know which coordinate system we are in (or have transformed to) before making these point-wise comparisons.
-In general, non-Cartesian or non-inertial coordinate systems will not allow for point-wise comparisons of the density.
-The Hamiltonian formalism circumvents this further epistemic requirement by defining the density using quantities that vary covariantly.
-Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
-
-[Should we briefly illustrate or gesture at an illustration of a failure of point-wise comparison?
-E.g., consider a situation where the density is a constant $k$ as a function of Cartesian coordinates (x, y).
-If we transform to polar coordinates $(r,\theta) $, then the scalar value of the density is no longer a constant (due to the Jacobian of this transformation not being unity).
-For instance, the scalar value of the density increases as we go further out, even though the density is physically the same.
-Hence, to determine whether there is more stuff at one location in state space vs.
-another, we need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
-we can't just compare the values of $\rho$ in these coordinates and see which is larger).]
-
-% Connect this back with fundamentality: epistemic notion of the fundamental (don't need to know which coordinate system we are in), along with a physical sense: supporting a physical invariant, namely the invariant phase space density.
-Showing that this is more physically significant than the state variables themselves.
+\subsection{Coordinate-dependent yet Fundamental}
+\label{fundamental}
+
+\begin{comment}
+
+Section on Ham. Not everything physical has to be coordinate independent or invariant.
+
+* In phase space, we do have coordinate invariance and point-wise comparison.
+Momentum changes covariantly.
+It is the density that is truly fundamental, not the coordinates in general.
+Though, the q's are priviledged from epistemic access.
+The p's are not, in fact we can define gauge transformation as things that keep the q's the same but change the p.
+
+*[Incorporate the following from section 5.2] Not all ``state variables" are physical.
+Hamiltonian mechanics does not really care what the q's and p's are.
+It is the motion assumption that tells us that q's are position, simply because they define the unit system.
+P is a gauge quantity (we care about the trajectory, the velocity).
+In this view, p is simply a ``label'' we can use to express the density over states.
+Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves.
+$->$ Invariant density is really the important physical element/object in Hamiltonian mechanics.
+
+* Units are physically significant.
+In fact, the ability to express the densities in units is what gives us the symplectic structure.
+It's a physical requirement that we must be able to express the density in some unit.
+* Invariant densities: physical significance connected to coordinate invariance. [incorporated]
+* Density comparison is independent of coordinates. We can compare densities point-wise without worry. [incorporated]
+
+* Makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
+Areas/volumes are measures of number of states.
+They are coordinate invariant.
+Also the Shannon/Gibbs entropy are coordinate invariant.
+
+% Section~\ref{action} uses the invariant density to provide a geometrical interpretation of the action principle.
+
+\end{comment}
+
+% Goal of the following paragraph is to explain the various senses in which the invariant phase space density is fundamental
 
@@
 We argued that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in part because it is more general.
-By considering how Hamiltonian mechanics supports faithful point-wise density comparisons, we have seen two further notions of fundamentality.
+By considering how Hamiltonian mechanics supports point-wise density comparisons, we have encountered two further notions of fundamentality.
 The first is epistemic: by representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
@@
 If we did not define the density, then we would need another explanation of why this mathematical structure arises for classical systems.
-[I think this final point might be repetitive given a claim made in the next paragraph]
 %see 11:50am of 12-21 meeting video for clarification of this point. There are probably finer details for me to incorporate.
 
-% We took out the following because coordinate-invariance in fact did not seem to be either a necessary or sufficient condition for physical significance: Assuming that coordinate-invariance is sufficient for physical significance, the density thereby represents a genuine physical property of classical systems.
-
-[To do: expand on the following point: ] This is why statistical mechanics works best on phase space (we have a uniform measure over q/p). See the relevant * point above.
-
-% Point of the following paragraph is to draw a contrast with North, who also seeks to identify physically significant quantities in terms of coordinate-invariants.
-We are pointing out that coordinate-invariance is not (on our view) a necessary condition for physical significance.
-
-While it is important to provide a physical interpretation of invariants, this is not to say that only invariant quantities are physically meaningful.
-For instance, the coordinates $q$ define the units, and units are physically significant despite not being invariant.
-Units are physically significant because they allow us to assign numbers to a physical quantity.
-They are physically significant in part because they correspond to measurable quantities.
-Indeed, as the derivation in Section~\ref{deterministic} shows, a necessary condition for defining the symplectic structure is that we can express the density in a particular choice of units.
-If we did not have the physical ability to define the density in some units---although any particular choice is arbitrary---we would not be able to define this structure.
-[To do: make this argument more clear]
-
-Whereas North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}, some choices of the coordinates $q$ are less physically meaningful because they are not properly connected with measurement.
-Some arbitrary mixings of $q$'s and $p$'s preserve the symplectic manifold but are not as physically significant.
-For this reason, the $q$'s are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
-Whereas symplectic geometry creates a mathematical, syntactic symmetry between the $q$ and $p$ variables, this is not a physically significant symmetry from the standpoint of measurement.
-%cite van Fraassen on the measurement point; e.g. APA session on Halvorson book
-
-Further support for the preceding point: could discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
-(In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).
-
-% Have mostly done the following: [To do: expand on the following point, and try to eliminate the notion of ``simplicity" and replace it with a more epistemic notion.
-Ability to make comparisons of density, independently of which coordinate system you are in or knowledge of being in a ``good" coordinate system.
-Josh: see 11-30-20 meeting notes where I think we discussed this.]: Compared with Lagrangian mechanics, canonical pairs (position and conjugate momentum) allow a simpler comparison of densities over space.
-
-[Not sure if/where the following claim is relevant: ] This difference in the number of assumptions corresponds to a difference in the number of commitments.
-When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?).
-In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.
-
-% By definition, a dynamical system is one where we can write xdot as a function of x (i.e.
-the current state determines the evolution of the system).
-Having a Hamiltonian system further means that we can express this function xdot(x) using the Hamiltonian.
-
-\section{Interpreting Lagrangian mechanics}
-\label{Lagrangian}
-
-By analyzing classical systems in terms of their invariant phase space density, we gain new insights into Lagrangian mechanics.
-Section~\ref{action} shows how the action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
-We show that the action principle holds even when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
-Section~\ref{gauge} discusses how the Lagrangian is highly gauge-redundant, and therefore less physical than the Hamiltonian.
-This provides a further argument for privileging Hamiltonian mechanics in the interpretation of classical systems.
-Finally, Section~\ref{Curiel} considers Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics, which he alleges encodes the necessary and sufficient structure for capturing the physically significant kinematical constraints of classical systems.
-This raises a number of objections to our own position, which we rebut.
-In short, we argue that we have given an eminently natural construction of Lagrangian mechanics, viewed as a special case of Hamiltonian mechanics.
-Moreover, our argument shows that Lagrangian mechanics is in some sense less natural than Hamiltonian mechanics because the former is contingent on having a bijection between velocity and conjugate momentum.
-
-\subsection{A geometric interpretation of the action principle}
-\label{action}
-
-% Somewhere below, emphasize that you have a unique trajectory only if the action is a minimum or maximum (if it is just stationary but not a minimum or maximum, then you do not have a unique trajectory)
-
-Hamilton's principle of stationary action serves as a foundational principle for Lagrangian mechanics.
-According to this principle, infinitesimal variations of the action $ \mathscr{A}$ along the system's trajectory $\gamma $ equal zero: $\delta  \mathscr{A}[\gamma] = 0 $.
-Although central to Lagrangian mechanics, we can actually derive the stationary action principle using Hamiltonian mechanics.
-Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition.
-We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating the generalization, see \textcites[110]{Arnold1989}.
-[To do: verify that Arnold can legitimately be read as generalizing the argument below.]}
-% Note on the generalization: cannot use curl.
-Need to use a form that measures the number of states on a surface.
-The displacement field kills this form (eats the degenerate direction that does not form degrees of freedom).
-% Something that I thought was a second assumption but turns out that this follows just from double role played by time variable: 2) the system's motion always advances in time (so that $dt/dt = 1$).
-
-Knowing that the system is deterministic and reversible, it follows that the density of states is invariant.
-This means that possible trajectories of the system are never pushed toward each other or pulled apart.
-In other words, the flux of trajectories through any closed surface is always zero.
-Formally, the displacement of the states in time is divergenceless: $\nabla \cdot \vector{S} = 0 $.
-Here, $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ defines a vector field in the extended phase space $(q, p , t ) $.
-$\vector{S}$ characterizes the displacement of states in time.
-In particular, it is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and affine parameter of the trajectory.
-
-Since $\vector{S}$ is divergenceless, there exists another vector field $-\vector{\theta}$ such that $\vector{S} = curl(-\vector{\theta} )$.\footnote{For reasons that will become clear, we choose $-\vector{\theta}$ rather than $\vector{\theta}$.} As a vector field in the extended phase space, $\vector{\theta}$ can be written in terms of its components $\vector{\theta} = (\theta_q, \theta_p, \theta_t) $.
-Since the curl of the gradient of a scalar function is always zero, the displacement vector $\vector{S}$ is invariant under the gauge transformation $\vector{\theta'} =\vector{\theta} + grad(f)$, where $f$ is an arbitrary scalar function.
-Hence, we can choose $f$ such that $\partial f/ \partial p = -\theta_p $.
-Without loss of generality, we can thereby redefine $\vector{\theta}$ as $(\theta_q, 0, \theta_t) $.
-
-Treating time $t$ as both a coordinate and an affine parameter (parameterizing the trajectory), the time component of the displacement, $S_t = dt/ dt = 1$.\footnote{Considering $dt/dt$, the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
-We could instead use a parameter $\tau$, and the derivation would remain valid.} This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $.
-In components, this means that $S_t = - (\frac{\partial}{\partial q} \cdot \theta_p - \frac{\partial}{\partial p} \cdot \theta_q) = - (\frac{\partial}{\partial q} \cdot 0 - \frac{\partial}{\partial p} \cdot \theta_q) = \frac{\partial q}{\partial p}$.
-Hence, $S_t = 1$ entails that $\frac{\partial q}{\partial p} = 1$.
-Integrating this expression, we find that $\theta_q = p$ up to addition by an arbitrary constant function $g(q)$.
-Again using the gauge freedom of $\vector{S}$ we can set $g(q)=0$, so that $\theta_q = p$.
-Thus, so far, we have shown that $\vector{\theta} = (p, 0,\theta_t) $.
-
-Next, we introduce a function $H$ and let $ \theta_t \equiv -H$.
-Hence, we write $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
-By matching components of $\vector{S}$ and $-curl(\theta)$, we show that H is the Hamiltonian for the system.
-First, by definition, $S_q \equiv \frac{dq}{dt} $ and $S_p \equiv  \frac{dp}{dt} $.
-Secondly, since $\vector{S} = -curl(\vector{\theta}) $, we calculate that $S_q = -\frac{\partial \theta_t}{\partial p} = \frac{\partial H}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}=-\frac{\partial H}{\partial q}$.
-Matching terms, we see that $\frac{\partial H}{\partial p} =\frac{dq}{dt} $ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt}$.
-These simply are Hamilton's equations, which justifies interpreting $H$ as the Hamiltonian.
-
-% Since the system is deterministic and reversible, it satisfies Hamilton's equations: $\frac{\partial H}{\partial p} =\frac{dq}{dt} \equiv S_q$ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt} \equiv S_p $.
-Additionally, since $\vector{S} = -curl(\vector{\theta}) $, we know that $S_q = -\frac{\partial \theta_t}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}$.
-Integrating the first equation for $\theta_t$ gives $-H$ up to an arbitrary constant function of q, while the second gives $\theta_t = -H$ up to an arbitrary constant function of p.
-For consistency, these two constant functions must vanish, so that $\theta_t = -H$ [check argument here; do we actually need to appeal to the gauge freedom of S again?].
-Hence, it follows that we can take $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
-
-At this point, we return to the action $ \mathscr{A}$, defined as the integral of the Lagrangian over a trajectory beginning at time $t_0$ and ending at time $t_1$: $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt$.
-Provided that the Legendre transform is well-defined (i.e.
-the motion equivalence assumption is satisfied), then $H = p \dot{q} - L$.
-Hence, we can re-express the action as $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$.
-Multiplying the Hamiltonian by $dt/dt = 1$ and adding a vanishing $0 \cdot \frac{dp}{dt}$, we obtain $ \mathscr{A} = \integral^{t_1}_{t_0} [p \frac{dq}{dt} + 0 \cdot \frac{dp}{dt} - H(q, p, t) \frac{dt}{dt} ]dt$.
-We notice that the integrand simply equals the inner product of $\vector{\theta}$ with $\vector{S} $, entailing that the Lagrangian $L =\vector{\theta} \cdot \vector{S} $.
-Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$.
-Changing variables, we take the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.\footnote{Note that $\vector{S}  dt = d\vector{\gamma}$ only on the actual trajectory in phase space, but not on other possible trajectories.} To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
-% Note on the generality of the argument: in cases where motion equivalence is not satisfied, we can still write the integrand in terms of H(q, p, t), although it would not be an integral over the Lagrangian.
-Thus, we can still express the `action' as a function of p, q, t.
-We would calculate this on a path in the extended phase space, although without motion equivalence, not all of the paths on phase space would correspond to a unique trajectory in real, physical space.
-See if Arnold has a more general name for this generalized principle of stationary action.
-
-By definition, a variation over the integral $\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ consists in taking the difference between the integrand and a corresponding integral over a nearby trajectory $\gamma'$: $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} - \integral_{\gamma'} \vector{\theta} \cdot d\vector{\gamma'}$ .
-Since the trajectories $\gamma$ and $\gamma'$ have the same endpoints, they define a closed-curve $C$, enclosing a region $\Sigma $.
-Hence, by Stoke's theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot dC = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ .
-Recalling now that 1) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and 2)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $\vector{\theta}$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
-This establishes Hamilton's principle of stationary action, which we have derived by assuming that 1) the displacement vector is divergenceless and 2) using the Legendre transform (assuming motion equivalence).
-
-(XXX: Note how the geometry is better understood in extended phase space, requiring Hamiltonian picture.
-Note that this applies even when $\theta S$ cannot be expressed as function of $q$ and $\dot q$.
-)
-
-%Discussion of gauge transformations in the context of Lagrangian vs.
-Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):
-
-\subsection{Gauge redundancy in Lagrangian mechanics}
-\label{gauge}
-
-\begin{comment}
-
-% 12-19: I think that I incorporated the relevant points below:
-
-** Overall point: this approach allows us to get a better intuition for what gauge transformations represent.
-
-* From motion equivalence -> we really only care about the trajectory. That is what we want to represent and what we measure.
-
-* A gauge transformation is a change in p variable that leave q and q dot unchanged.
-Therefore it is simply a change of how states are labeled, without a change in the actual observable dynamics.
-Add the derivation.
-
-* In Hamiltonian mechanics, the momentum is gauge dependent.
-In Lagrangian mechanics the kinematic variables are not gauge dependent, but the Lagrangian manifestly is.
-$L = \theta S$.
-$S$ is gauge independent but $\theta$ is not.
-Therefore we will have problems understanding the physicality of the Lagrangian.
-Similar concerns for the Action.
--> neither formulation has the upper hand
-
-* Coordinate transformation: if we restrict to coordinate transformation that only depend on space, both Lagrangian and Hamiltonian are invariant.
-If we introduce coordinate transformation that do depend on time (e.g.
-rotating frame), then Lagrangian is still invariant while Hamiltonian manifestly changes as a time component of a co-vector.
-(because H is the time component of $\theta$).
-That is why the Lagrangian framework is prevalently used in intro classes.
-
-* This is a problem of the double role of time as ``coordinate'' and ``affine parameter''.
-To be proper in differential geometry, we have to separate the two uses.
-So we add a ``temporal'' degree of freedom $(t, E)$ and the evolution is over the affine parameter $s$.
-In that framework, the Hamiltonian constraint serves both as a global constraint that fixes the energy as a function of $q, p, t$ and as the generator of the affine parameter $s$.
-The Hamiltonian constraint is invariant.
-If the affine parameter is proper time, then the Hamiltonian constraint is essentially the mass.
--> There is a better formulation of Hamiltonian mechanics, that is typically not found in textbooks, that addresses the problem.
-
-\end{comment}
-
-% Last page of Halvorson/Swanson article seems to broach this issue of gauge redundancy, although they seem to think it is possibly a positive feature of Lagrangian mechanics (a source of additional representational flexibility) rather than a defect.
-
-Understanding both mechanics in terms of our assumptions gives us a way to understand gauge transformations that is both more physically grounded and intuitive.
-Recall that motion equivalence entails that knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
+Collectively, these considerations speak in favor of viewing the invariant phase space density as not only physically significant but also fundamental.
+Yet, as Section~\ref{dependence} explained, the density is always coordinate-dependent, even when coordinate-invariant.
+In other words, there is no coordinate-independent (or coordinate-free definition) of the invariant phase space density.
+Affording the density fundamental significance therefore places us at odds with those who take coordinate-independence as a necessary condition for fundamentality.
+[To do: provide literature references for this and motivating quotations]
+
+Some philosophers have instead privileged the state variables $(q, p)$.
+For instance, in her argument for the fundamentality of Hamiltonian mechanics, North claims that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
+[To do: see if Wallace 2003 also makes this claim about the fundamentality of conjugate momentum, or the similarly flawed claim that conjugate momentum and position are on equal footing within Hamiltonian mechanics].
+Yet, as we show below, the conjugate momentum $p$ is actually a gauge-redundant quantity.
+Assuming that gauge-invariance is a necessary condition for fundamental physical significance, this precludes $p$ from having fundamental status.
+Instead, we should understand the conjugate momentum $p$ as a label that we can use to express the invariant density over states.
+In this way, the invariant density is more physical than the state variables themselves.
+% Possibly a stronger claim: This entails that the invariant density is really the important physical element/object in Hamiltonian mechanics.
+
+[To do: explain why we have grounds to privilege the invariant density over not only the conjugate momenta but also the canonical position coordinate.
+Is this connected with the following thought: Hamiltonian mechanics does not really care what the q's and p's are.
+It is the motion assumption that tells us that q's are position, simply because they define the unit system.
+]
+
+[To do: think about whether treating gauge-invariance as a necessary condition for fundamentality is in tension with our position, specifically our contention that coordinate-dependency does NOT preclude a quantity from fundamental physical significance.
+Does our position require an important physical difference between gauge-redundancy and coordinate-dependence?].
+% % Note that if coherent, our position has the following consequence: gauge-redundancy is a sufficient condition for a quantity to be non-fundamental.
+But coordinate-dependency is NOT a sufficient condition for being non-fundamental.
+This could strike some readers as in tension.
+If we are comfortable viewing gauge-redundancy as precluding a quantity from fundamental significance, then why not coordinate-dependence?
+
+We now explain why the conjugate momentum $p$ is gauge-dependent.
+Recall that motion equivalence entails the following: knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
 In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$.
 Hence, any transformation that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant.
-We want such transformation to be canonical as to preserve the invariant densities.
-This means that the new state variable $(q, \hat{p})$ will respect the usual Poisson brackets:
+We want such transformations to be canonical in order to preserve the invariant density.
+This means that the new state variables $(q, \hat{p})$ will respect the usual Poisson brackets:
 \begin{equation}
@@
 Therefore changes of state variable of the form $\hat{p}_j = p_j + \partial_j f$ do not change either the equations of motion or the trajectories.
-We recognize these as the gauge transformations.
-
-Understanding the action principle in terms of Hamiltonian mechanics also provides a better understanding of gauge transformations.
-If motion equivalence holds, then gauge transformations leave the variables $(q,\dot{q}) $ invariant.
-To see this, recall that motion equivalence entails that knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
-In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$.
-Hence, any transformation that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant, i.e.
-is a gauge transformation.
-[To do: make it more clear how the discussion in this section is intrinsically connected with the earlier geometric treatment of the action and our overall privileging of Hamiltonian mechanics]
-
-In particular, there is a family of gauge transformations that changes the conjugate momentum variable $p$ while leaving $(q,\dot{q}) $ invariant.
-[To do: illustrate this with general formula for this kind of gauge transformation, and possibly the electrodynamics example.] Consequently, the conjugate momentum in Hamiltonian mechanics is gauge-dependent.
-The $p $ variables are therefore not of fundamental physical significance [to do: check that this is consistent with Section 2.1's discussion of fundamentality].
-Hence, privileging Hamiltonian mechanics does not entail, \textit{pace} North, that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
-[To do: see if Wallace 2003 also makes this claim about the fundamentality of conjugate momentum, or the similarly flawed claim that conjugate momentum and position are on equal footing within Hamiltonian mechanics].
+We recognize these as a family of gauge transformations that change the conjugate momentum variable $p$ while leaving $(q,\dot{q}) $ invariant.
+The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
+Additionally, it shows that not all state variables are on equal physical footing.
 Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.
-
-Instead, position and kinetic momentum are physically privileged, since these quantities are gauge-invariant.
+Indeed, as we now argue, some coordinate choices are more significant than others.
+% Whereas symplectic geometry creates a mathematical, syntactic symmetry between the $q$ and $p$ variables, this is not a physically significant symmetry from the standpoint of measurement.
+%cite van Fraassen on the measurement point; e.g. APA session on Halvorson book
+
+In privileging the phase space density, we have relied on its coordinate-invariance.
+Plausibly, coordinate-invariance is a necessary condition for fundamental physical significance.
+Being necessary for fundamental significance does not, however, make coordinate-invariance necessary for physical significance simpliciter.
+Our position here again places us at odds with others in the literature.
+For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
+Yet, some choices of the coordinates $q$ are not appropriately connected with physical measurement, making them less physically meaningful [to do: provide a clear example of this].
+Crucially, the coordinates $q$ define the units of physical measurement, allowing us to assign numbers to physical quantities.
+These units are physically significant despite not being coordinate-invariant.
+Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are physically significant.
+In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative  $p$'s are not.
+
+[Possible further point to include: discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
+(In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).]
+
+[Not sure if/where to include the following point (possibly with the discussion of the coordinate-dependence of the density?): As the derivation in Section~\ref{deterministic} shows, a necessary condition for defining the symplectic structure is that we can express the density in a particular choice of units.
+If we did not have the physical ability to define the density in some units we would not be able to define this structure.
+]
+
+[To do: expand on the following point: ] This is why statistical mechanics works best on phase space (we have a uniform measure over q/p).
+Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
+Areas/volumes are measures of number of states.
+They are coordinate invariant.
+Also the Shannon/Gibbs entropy are coordinate invariant.
+
+% Have mostly done the following: [To do: expand on the following point, and try to eliminate the notion of ``simplicity" and replace it with a more epistemic notion.
+Ability to make comparisons of density, independently of which coordinate system you are in or knowledge of being in a ``good" coordinate system.
+Josh: see 11-30-20 meeting notes where I think we discussed this.]: Compared with Lagrangian mechanics, canonical pairs (position and conjugate momentum) allow a simpler comparison of densities over space.
+
+% [Not sure if/where the following claim is relevant: ] This difference in the number of assumptions corresponds to a difference in the number of commitments.
+When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?).
+In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.
+It is the motion equivalence assumption that tells us that q's are position, simply because they define the unit system.
+P is a gauge quantity (we care about the trajectory, the velocity).
+
+% By definition, a dynamical system is one where we can write xdot as a function of x (i.e.
+the current state determines the evolution of the system).
+Having a Hamiltonian system further means that we can express this function xdot(x) using the Hamiltonian.
+
+% We took out the following because coordinate-invariance in fact did not seem to be either a necessary or sufficient condition for physical significance: Assuming that coordinate-invariance is sufficient for physical significance, the density thereby represents a genuine physical property of classical systems.
+
+\section{Interpreting Lagrangian mechanics}
+\label{Lagrangian}
+
+By analyzing classical systems in terms of their invariant phase space density, we gain new insights into Lagrangian mechanics.
+Section~\ref{action} shows how the action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
+We show that the action principle holds even when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
+Section~\ref{gauge} discusses how the Lagrangian is highly gauge-redundant, and therefore less physical than the Hamiltonian.
+This provides a further argument for privileging Hamiltonian mechanics in the interpretation of classical systems.
+Finally, Section~\ref{Curiel} considers Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics, which he alleges encodes the necessary and sufficient structure for capturing the physically significant kinematical constraints of classical systems.
+This raises a number of objections to our own position, which we rebut.
+In short, we argue that we have given an eminently natural construction of Lagrangian mechanics, viewed as a special case of Hamiltonian mechanics.
+Moreover, our argument shows that Lagrangian mechanics is in some sense less natural than Hamiltonian mechanics because the former is contingent on having a bijection between velocity and conjugate momentum.
+
+\subsection{A geometric interpretation of the action principle}
+\label{action}
+
+% Somewhere below, emphasize that you have a unique trajectory only if the action is a minimum or maximum (if it is just stationary but not a minimum or maximum, then you do not have a unique trajectory)
+
+Hamilton's principle of stationary action serves as a foundational principle for Lagrangian mechanics.
+According to this principle, infinitesimal variations of the action $ \mathscr{A}$ along the system's trajectory $\gamma $ equal zero: $\delta  \mathscr{A}[\gamma] = 0 $.
+Although central to Lagrangian mechanics, we can actually derive the stationary action principle using Hamiltonian mechanics.
+Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition.
+We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating the generalization, see \textcites[110]{Arnold1989}.
+[To do: verify that Arnold can legitimately be read as generalizing the argument below.]}
+% Note on the generalization: cannot use curl.
+Need to use a form that measures the number of states on a surface.
+The displacement field kills this form (eats the degenerate direction that does not form degrees of freedom).
+% Something that I thought was a second assumption but turns out that this follows just from double role played by time variable: 2) the system's motion always advances in time (so that $dt/dt = 1$).
+
+Knowing that the system is deterministic and reversible, it follows that the density of states is invariant.
+This means that possible trajectories of the system are never pushed toward each other or pulled apart.
+In other words, the flux of trajectories through any closed surface is always zero.
+Formally, the displacement of the states in time is divergenceless: $\nabla \cdot \vector{S} = 0 $.
+Here, $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ defines a vector field in the extended phase space $(q, p , t ) $.
+$\vector{S}$ characterizes the displacement of states in time.
+In particular, it is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and affine parameter of the trajectory.
+
+Since $\vector{S}$ is divergenceless, there exists another vector field $-\vector{\theta}$ such that $\vector{S} = curl(-\vector{\theta} )$.\footnote{For reasons that will become clear, we choose $-\vector{\theta}$ rather than $\vector{\theta}$.} As a vector field in the extended phase space, $\vector{\theta}$ can be written in terms of its components $\vector{\theta} = (\theta_q, \theta_p, \theta_t) $.
+Since the curl of the gradient of a scalar function is always zero, the displacement vector $\vector{S}$ is invariant under the gauge transformation $\vector{\theta'} =\vector{\theta} + grad(f)$, where $f$ is an arbitrary scalar function.
+Hence, we can choose $f$ such that $\partial f/ \partial p = -\theta_p $.
+Without loss of generality, we can thereby redefine $\vector{\theta}$ as $(\theta_q, 0, \theta_t) $.
+
+Treating time $t$ as both a coordinate and an affine parameter (parameterizing the trajectory), the time component of the displacement, $S_t = dt/ dt = 1$.\footnote{Considering $dt/dt$, the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
+We could instead use a parameter $\tau$, and the derivation would remain valid.} This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $.
+In components, this means that $S_t = - (\frac{\partial}{\partial q} \cdot \theta_p - \frac{\partial}{\partial p} \cdot \theta_q) = - (\frac{\partial}{\partial q} \cdot 0 - \frac{\partial}{\partial p} \cdot \theta_q) = \frac{\partial q}{\partial p}$.
+Hence, $S_t = 1$ entails that $\frac{\partial q}{\partial p} = 1$.
+Integrating this expression, we find that $\theta_q = p$ up to addition by an arbitrary constant function $g(q)$.
+Again using the gauge freedom of $\vector{S}$ we can set $g(q)=0$, so that $\theta_q = p$.
+Thus, so far, we have shown that $\vector{\theta} = (p, 0,\theta_t) $.
+
+Next, we introduce a function $H$ and let $ \theta_t \equiv -H$.
+Hence, we write $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
+By matching components of $\vector{S}$ and $-curl(\theta)$, we show that H is the Hamiltonian for the system.
+First, by definition, $S_q \equiv \frac{dq}{dt} $ and $S_p \equiv  \frac{dp}{dt} $.
+Secondly, since $\vector{S} = -curl(\vector{\theta}) $, we calculate that $S_q = -\frac{\partial \theta_t}{\partial p} = \frac{\partial H}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}=-\frac{\partial H}{\partial q}$.
+Matching terms, we see that $\frac{\partial H}{\partial p} =\frac{dq}{dt} $ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt}$.
+These simply are Hamilton's equations, which justifies interpreting $H$ as the Hamiltonian.
+
+% Since the system is deterministic and reversible, it satisfies Hamilton's equations: $\frac{\partial H}{\partial p} =\frac{dq}{dt} \equiv S_q$ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt} \equiv S_p $.
+Additionally, since $\vector{S} = -curl(\vector{\theta}) $, we know that $S_q = -\frac{\partial \theta_t}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}$.
+Integrating the first equation for $\theta_t$ gives $-H$ up to an arbitrary constant function of q, while the second gives $\theta_t = -H$ up to an arbitrary constant function of p.
+For consistency, these two constant functions must vanish, so that $\theta_t = -H$ [check argument here; do we actually need to appeal to the gauge freedom of S again?].
+Hence, it follows that we can take $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
+
+At this point, we return to the action $ \mathscr{A}$, defined as the integral of the Lagrangian over a trajectory beginning at time $t_0$ and ending at time $t_1$: $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt$.
+Provided that the Legendre transform is well-defined (i.e.
+the motion equivalence assumption is satisfied), then $H = p \dot{q} - L$.
+Hence, we can re-express the action as $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$.
+Multiplying the Hamiltonian by $dt/dt = 1$ and adding a vanishing $0 \cdot \frac{dp}{dt}$, we obtain $ \mathscr{A} = \integral^{t_1}_{t_0} [p \frac{dq}{dt} + 0 \cdot \frac{dp}{dt} - H(q, p, t) \frac{dt}{dt} ]dt$.
+We notice that the integrand simply equals the inner product of $\vector{\theta}$ with $\vector{S} $, entailing that the Lagrangian $L =\vector{\theta} \cdot \vector{S} $.
+Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$.
+Changing variables, we take the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.\footnote{Note that $\vector{S}  dt = d\vector{\gamma}$ only on the actual trajectory in phase space, but not on other possible trajectories.} To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
+% Note on the generality of the argument: in cases where motion equivalence is not satisfied, we can still write the integrand in terms of H(q, p, t), although it would not be an integral over the Lagrangian.
+Thus, we can still express the `action' as a function of p, q, t.
+We would calculate this on a path in the extended phase space, although without motion equivalence, not all of the paths on phase space would correspond to a unique trajectory in real, physical space.
+See if Arnold has a more general name for this generalized principle of stationary action.
+
+By definition, a variation over the integral $\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ consists in taking the difference between the integrand and a corresponding integral over a nearby trajectory $\gamma'$: $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} - \integral_{\gamma'} \vector{\theta} \cdot d\vector{\gamma'}$ .
+Since the trajectories $\gamma$ and $\gamma'$ have the same endpoints, they define a closed-curve $C$, enclosing a region $\Sigma $.
+Hence, by Stoke's theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot dC = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ .
+Recalling now that 1) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and 2)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $\vector{\theta}$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
+This establishes Hamilton's principle of stationary action, which we have derived by assuming that 1) the displacement vector is divergenceless and 2) using the Legendre transform (assuming motion equivalence).
+
+(XXX: Note how the geometry is better understood in extended phase space, requiring Hamiltonian picture.
+Note that this applies even when $\theta S$ cannot be expressed as function of $q$ and $\dot q$.
+)
+
+%Discussion of gauge transformations in the context of Lagrangian vs.
+Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):
+
+\subsection{Gauge redundancy in Lagrangian mechanics}
+\label{gauge}
+
+\begin{comment}
+
+% 12-19: I think that I incorporated the relevant points below:
+
+** Overall point: this approach allows us to get a better intuition for what gauge transformations represent.
+
+* From motion equivalence -> we really only care about the trajectory. That is what we want to represent and what we measure.
+
+* A gauge transformation is a change in p variable that leave q and q dot unchanged.
+Therefore it is simply a change of how states are labeled, without a change in the actual observable dynamics.
+Add the derivation.
+
+* In Hamiltonian mechanics, the momentum is gauge dependent.
+In Lagrangian mechanics the kinematic variables are not gauge dependent, but the Lagrangian manifestly is.
+$L = \theta S$.
+$S$ is gauge independent but $\theta$ is not.
+Therefore we will have problems understanding the physicality of the Lagrangian.
+Similar concerns for the Action.
+-> neither formulation has the upper hand
+
+* Coordinate transformation: if we restrict to coordinate transformation that only depend on space, both Lagrangian and Hamiltonian are invariant.
+If we introduce coordinate transformation that do depend on time (e.g.
+rotating frame), then Lagrangian is still invariant while Hamiltonian manifestly changes as a time component of a co-vector.
+(because H is the time component of $\theta$).
+That is why the Lagrangian framework is prevalently used in intro classes.
+
+* This is a problem of the double role of time as ``coordinate'' and ``affine parameter''.
+To be proper in differential geometry, we have to separate the two uses.
+So we add a ``temporal'' degree of freedom $(t, E)$ and the evolution is over the affine parameter $s$.
+In that framework, the Hamiltonian constraint serves both as a global constraint that fixes the energy as a function of $q, p, t$ and as the generator of the affine parameter $s$.
+The Hamiltonian constraint is invariant.
+If the affine parameter is proper time, then the Hamiltonian constraint is essentially the mass.
+-> There is a better formulation of Hamiltonian mechanics, that is typically not found in textbooks, that addresses the problem.
+
+\end{comment}
+
+% Last page of Halvorson/Swanson article seems to broach this issue of gauge redundancy, although they seem to think it is possibly a positive feature of Lagrangian mechanics (a source of additional representational flexibility) rather than a defect.
+
+% Understanding mechanics in terms of our assumptions gives us a way to understand gauge transformations that is both more physically grounded and intuitive.
+
+As discussed in Section~\ref{fundamental}, the conjugate momentum $p$ is a gauge-dependent variable.
+In contrast, position and kinetic momentum are both gauge-invariant, making them physically privileged.
 Consequently, Lagrangian mechanics employs state space variables that are gauge-invariant, while Hamiltonian mechanics uses a gauge-dependent variable.
@@
 In Section~\ref{Curiel}, we turn to considering Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.
-
-[Could address the following in this section, rather than in Section 4:] Not all ``state variables" are physical.
-Hamiltonian mechanics does not really care what the q's and p's are.
-It is the motion assumption that tells us that q's are position, simply because they define the unit system.
-P is a gauge quantity (we care about the trajectory, the velocity).
-In this view, p is simply a ``label'' we can use to express the density over states.
-Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves.
-This entails that the invariant density is really the important physical element/object in Hamiltonian mechanics.
 

### 2021-01-11 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Adding notes after meeting

@@
 While this provides one reason for privileging Hamiltonian mechanics, it is not the only one.
-A special relationship exists The special relationship between Hamiltonian mechanics and invariant densities provides a further---no less important---reason for privileging Hamiltonian mechanics.
+The special relationship between Hamiltonian mechanics and invariant densities provides a further---no less important---reason for privileging Hamiltonian mechanics.
 In this section, we argue that viewing physics through the lens of invariant densities has much to commend it.
@@
 In differential geometry, a different mathematical object is called the `density,' namely a volume form (i.e.
-an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation.
-Our physically significant density $\rho$ is actually the component of this volume form $\nu = \rho(x^i) dx^1 \wedge dx^2 \wedge ...
-\wedge dx^n$.
+an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation.\footnote{TODO: The differential density can be understood as a way to generalize volume forms in the case of a non-orientable manifold.
+This case is not interesting to us, therefore we will be glossing over issues of orientation and subtle differences between volume forms and differential densities.
+We believe this simplifies the exposition removing details that have no real consequence for the physical understanding.} Our physically significant density $\rho$ is actually the component of this volume form $\nu = \rho(x^i)  \epsilon dx^1 \wedge dx^2 \wedge ...
+\wedge dx^n$ where $\epsilon$ is $1$ or $-1$ depending on the orientation..
 Note that $\nu$ combines the density and the volume into a coordinate-free object.
@@
 In other words, $\nu$ has dimensions of only mass and is defined as a function of both a point and a volume element.
-In contrast, the density $\rho$ is defined at each point and has dimensions of mass over volume.
+In contrast, the density $\rho$ is just a function of the coordinates at a point, not the volume element, and has dimensions of mass over volume.
 This shows that the coordinate-independent function $\nu$ cannot play the role of the coordinate-dependent density $\rho$.
@@
 In Riemannian geometry, we have a canonical volume form $dV = \epsilon \sqrt{|g_{ab}|} dx^1 \wedge dx^2 \wedge ...
-\wedge dx^n$ where $|g_{ab}|$ is the determinant of the metric tensor as expressed in local coordinates, and $\epsilon$ is $1$ or $-1$ depending on the orientation.
+\wedge dx^n$ where $|g_{ab}|$ is the determinant of the metric tensor as expressed in local coordinates.
 In component notation, we can express the volume form as $\sqrt{|g_{ab}|}\epsilon_{a_1 ...
@@
 At first glance, this expression misleadingly suggests that the density $\hat{\rho}$ is coordinate-independent: it is a scalar quantity with the correct dimensions of mass over volume.
-Nevertheless, $\hat{\rho}$ remains coordinate-dependent because it is a function of $|g_{ab}|$, which depends on the choice of coordinates and is, in general, a function of the point.
+[XXX] Nevertheless, $\hat{\rho}$ remains coordinate-dependent because it is a function of $|g_{ab}|$, which depends on the choice of coordinates and is, in general, a function of the point.
 However, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
@@
 contravariantly) as differentials in position.
-Consequently, the density $\rho(q,\dot{q})$ also varies contravariantly.
-These transformation properties preclude the density---in the Lagrangian formalism---from being coordinate-invariant in general.
-In detail, $\rho(x^i, v^j) = \sqrt{|g_{ab}|} \sqrt{|g_{ab}|} \rho(\hat{x}^i, \hat{v}^j) = |g_{ab}| \rho(\hat{x}^i, \hat{v}^j)$.
-The determinant of the metric tensor is therefore directly the volume of the tangent bundle.
+Since both position and velocity transform contravariantly, the density will transform according to $\rho(x^i, v^j) = \sqrt{|g_{ab}|} \sqrt{|g_{ab}|} \rho(\hat{x}^i, \hat{v}^j) = |g_{ab}| \rho(\hat{x}^i, \hat{v}^j)$.
+The canonical volume form for the tangent bundle becomes: $dV_{TM} = \epsilon \sqrt{|g_{ab}|} dx^1 \wedge dx^2 \wedge ...
+\wedge dx^n \bigwedge \epsilon \sqrt{|g_{ab}|} dv^1 \wedge dv^2 \wedge ...
+\wedge dv^n = |g_{ab}| \bigwedge dx^i \bigwedge dv^j$.
+The determinant of the metric tensor is therefore directly the component of the volume form of the tangent bundle.
 The relationship to the density as expressed in phase space is therefore the following:
 \begin{equation}
-	\rho(q^i, p_j) = |g_{ab}| \rho(x^i, v^j).
+	\rho(q^i, p_j) = m |g_{ab}| \rho(x^i, v^j).
 \end{equation}
@@
 % Josh: why is the determinant of the metric tensor equal to the volume of the tangent bundle in this case? [Discussed on 1-4 meeting, 11:56 a.m.]. Perhaps study discussion of this in the long paper.
+
+[XXX] These relationships also allow us now to better understand what the hyperregular case, and the massive particle in particular, correspond to.
+Note that, in general, the relationship between densities over the tangent and cotangent bundles is, as we saw:
+\begin{equation}
+\rho(q^i, p_j) = \left|\frac{\partial v^j}{\partial p_k}\right| \rho(x^i, v^j)  = \left|\frac{\partial H}{\partial p_k \partial p_j}\right| \rho(x^i, v^j).
+\end{equation}
+Note that $\frac{\partial H}{\partial p_k \partial p_j} = m g^{kj}$ exactly in the case of massive particles, where the Hamiltonian (and the Lagrangian) are quadratic.
+This is the case where the densities in the tangent bundle can be properly understood in terms of Riemmanian geometry.
+For the more general case, we note:
+\begin{equation}
+\begin{aligned}
+\delta^i_j = \frac{\partial v^i}{\partial v_j} = \frac{\partial v^i}{\partial p_k}\frac{\partial p_k}{\partial v_j} &= \frac{\partial}{\partial p_k}\left(\frac{\partial H}{\partial p_i}\right) \frac{\partial }{\partial v_j} \left( \frac{\partial L}{\partial v_k} \right) \\
+\frac{\partial L}{\partial v^k \partial v^j} &= \left[ \frac{\partial H}{\partial p_k \partial p_j} \right]^{-1}
+\end{aligned}
+\end{equation}
+which means the Hessian of the Hamiltonian is the inverse of the Hessian of the Lagrangian, and so are their determinants.
+The hyperregular case corresponds to the case where the determinant of the Hessian of the Lagrangian is non-zero, which means we can express the densities of states over position and velocity.
 
@@
 We now explain why the conjugate momentum $p$ is gauge-dependent.
-Recall that motion equivalence entails the following: knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
+[XXX: first we want to answer: what is a gauge transformation?
+What does it represent physically?] Recall that motion equivalence entails the following: knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
 In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$.
@@
 We recognize these as a family of gauge transformations that change the conjugate momentum variable $p$ while leaving $(q,\dot{q}) $ invariant.
-The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
+[XXX: make it clear that we found gauge transformations with our different definition] The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
 Additionally, it shows that not all state variables are on equal physical footing.
@@
 Indeed, as we now argue, some coordinate choices are more significant than others.
+[XXX: coordinate or state variables?]
 % Whereas symplectic geometry creates a mathematical, syntactic symmetry between the $q$ and $p$ variables, this is not a physically significant symmetry from the standpoint of measurement.

### 2021-01-12 Articles/2019-Entropy/Source/thermodynamics.tex
After review with Christine

@@
 This means that the count of evolutions per state can never decrease under a deterministic process and it is maximized at equilibrium.
-We show how this simple idea can be used as a straightforward foundation of thermodyanmics and how it relates to other notions of entropy in physics.
+We show how this simple idea can be used as a straightforward foundation for thermodynamics and how it relates to other notions of entropy in physics.
 
@@
 Rather then derive thermodynamics bottom-up from a particular microscopic model, we want to derive it top-down from a set of requirements on the dynamics of the system.
-The results we find are, then, independent on the microscopic dynamics in the sense that they can in principle be realized in different ways.
+The results we find are, then, independent of the microscopic dynamics in the sense that they can in principle be realized in different ways.
 This is both more in line with the aims of the founders of thermodynamics and justifies why thermodynamic concepts have been successful in fields beyond physics.
-Moreover, in our approach thermodynamics is found as a specialization of a more general model, which we believe can be used as a common foundation for the different types mechanics.
+Moreover, in our approach thermodynamics is found as a specialization of a more general model, which we believe can be used as a common foundation for the different types of mechanics.
 
@@
 Counting the possible evolutions plays a fundamental role, in much the same way that counting states plays a role in statistical mechanics.
-In a deterministic and reversible process, all evolutions that start in one state will end in the same state: the evolutions cannot merge or diverge.
-If the process is deterministic but not reversible, however, the same final state can be reached from different initial states, and the evolution will merge.
+In a deterministic and reversible process, there is one to one correspondence between initial and final states: the evolutions cannot merge or diverge.
+If the process is deterministic but not reversible, however, the same final state will be reached from different initial states, and the evolutions will merge.
 If the deterministic process reaches an equilibrium, there will be no more merging.
@@
 Therefore, in this paper, we will only give a conceptual overview leaving the formal details to other papers.
-This will leave the key physical insight is more directly accessible, and also result in a narrative that can be used directly in an introductory course of thermodynamics.
+This will leave the key physical insight more directly accessible, and also result in a narrative that can be used directly in an introductory course of thermodynamics.
 
@@
 Namely, a process tells us what can and cannot happen in a given set of circumstances.
-We also assume we have a measure $\mu$ that allows us to determine the size of a set of evolutions, to ``count'' evolutions.\footnote{The precise conceptual and mathematical definition for $\mu$ is source of a number of technical considerations which we (will) address in a separate work.
+We also assume we have a measure $\mu$ that allows us to determine the size of a set of evolutions, to ``count'' evolutions.\footnote{The precise conceptual and mathematical definition for $\mu$ is a source of a number of technical considerations which we (will) address in a separate work.
 While some issues, like unit definition, are important to the physics, most have to do with the proper handling of infinities or the conditions for existence and uniqueness of the measure.
@@
 \end{equation}
-Conceptually, the ability to count evolutions is equivalent to knowing all possible probability of transition.\footnote{Technically, a single probability measure is not enough to define all these transition at all scales, but the overall intuition remains.}
+Conceptually, the ability to count evolutions is equivalent to knowing all possible transition probabilities.\footnote{Technically, a single probability measure is not enough to define all these transitions at all scales, but the overall intuition remains.}
 
 Having answered the question of how we characterize processes, we can categorize them.
-For example, we say a process is \textbf{deterministic} if the state at one time in enough to determine the state at all future times.
+For example, we say a process is \textbf{deterministic} if the state at one time is enough to determine the state at all future times.
 We can see an example of such a process in figure \ref{fig_determinism}.
@@
 	\includegraphics[width=\columnwidth]{images/Slide3.png}
-	\caption{Deterministic process: the state one time determines the state at future times.
+	\caption{Deterministic process: the state at one time determines the state at future times.
 Evolutions can merge but not diverge in time.
@@
 
-%TODO: add a double evolution in some case.
 \begin{figure}[h!]
@@
 
-To finish, consider a \textbf{deterministic processes with equilibria}.
-We can see an example of such a process in figure \ref{fig_with_equilibria}.
-An equilibrium for a process is a state that, once it is reached, it is never left.\footnote{We will see later that this characterization of equilibrium is symplistic, but it will suffice for now.} In a process with equilibria all evolutions will reach, at some point, an equilibrium.
+To finish, consider a \textbf{deterministic processes with equilibria}, like the one in figure \ref{fig_with_equilibria}.
+An equilibrium for a process is a state that, once it is reached, it is never left.\footnote{We will see later that this characterization of equilibrium is simplistic, but it will suffice for now.} In a process with equilibria all evolutions will reach, at some point, an equilibrium.
 In the out-of-equilibrium phase, evolutions merge until they reach equilibria and can no longer merge.
@@
 Evolutions keep merging until equilibrium is reached.
-The number of evolutions per state increase and the maximum is reached at equilibria.}\label{fig_with_equilibria}
+The number of evolutions per state increases and the maximum is reached at equilibria.}\label{fig_with_equilibria}
 \end{figure}
 
-We already reached an important result: with only a handful of simple definitions we have found a quantity, $\mu(x(t))$, that can never decrease in a deterministic process, stays the same if the process is reversible and is maximized at equilibrium.
+We already reached an important result: with only a handful of simple definitions we have found a quantity, $\mu(x(t))$, that can never decrease in a deterministic process, stays the same if the process is also reversible and is maximized at equilibrium.
 We have done so with no assumption on the type of system, the type of forces acting on the system, the internal composition of the system or the internal dynamics of the system.
@@
 
-Suppose we now have two systems, with their respective state space $X$ and $Y$.
+Suppose we now have two systems, with their respective state spaces $X$ and $Y$.
 The state space of the composite system is the set of all possible pairs $X \times Y$.
@@
 \end{equation}
-the measure factorizes. We say two system are \textbf{independent} at given time $t$ if the evolutions per state factorizes: $\mu_{X \times Y}(x(t), y(t)) = \mu_X (x(t)) \mu_Y (y(t))$.
+the measure factorizes. We say two systems are \textbf{independent} at a given time $t$ if the evolutions per state factorizes: $\mu_{X \times Y}(x(t), y(t)) = \mu_X (x(t)) \mu_Y (y(t))$.
 
@@
 We define the \textbf{process entropy} as $S = \log(\mu)$.
-Since the logarithm is a monotonic function, the process entropy will preserve monotonicity and maximization of $\mu$ under deterministic process with equilibria.
-Which means:
+Since the logarithm is a monotonic function, the process entropy will preserve monotonicity and maximization of $\mu$ under a deterministic process with equilibria.
+This means:
 \begin{equation}\label{ov_entropy_increases}
@@
 In addition, the process entropy will be additive for independent systems.\footnote{As before there are a number of technical details that need to be addressed, most of all the units for $\mu$ which will influence the 0 point for $S$.
-We (will) address in the full framework.}
+We will address this in the full framework.}
 
@@
 If the system is a composite there are additional considerations.
-Different processes may put different degree of correlation between subsystems, therefore the same equilibria for the subsystems may correspond to different processes entropy values for the composite.
+Different processes may put different degrees of correlation between subsystems, therefore the same equilibria for the subsystems may correspond to different process entropy values for the composite.
 But if the subsystems are independent, the composite will correspond to a unique entropy value: the product.
 
-We can then assign a unique process entropy to states of equilibria and to composite state of independent equilibria.
-In these conditions we define the \textbf{state entropy} $S(x)$ which corresponds to the process entropy that the system reaches at equilibrium.
+We can then assign a unique process entropy to states of equilibria and to composite states of independent equilibria.
+Under these conditions we define the \textbf{state entropy} $S(x)$ which corresponds to the process entropy that the system reaches at equilibrium.
 But why would we want to assign a unique process entropy to states in the first place?
@@
 Suppose that we are studying a system and we change the circumstances surrounding it.
-It first evolves according to $A$ and we switch to process $B$.
+It first evolves according to $A$ and at $t_0$ we switch to process $B$.
 If $x(t)$ is an evolution of the combined process $AB$, then for $t\leq t_0$ it must be an evolution for $A$ and for $t\geq t_0$ it must be an evolution for $B$.
 Moreover, the number of total evolutions must remain the same.
-Therefore the state $x(t_0)$, the state where we are joining the evolutions, must have a well defined entropy that is independent on the process, or not all the evolutions would match.
+Therefore the state $x(t_0)$, the state where we are joining the evolutions, must have a well defined entropy that is independent of the process, or not all the evolutions would match.
 If we want to mix and match from a set of possible processes, then the process entropy at the junction must be only a function of the state.
 
-The second reason is that, upon close inspection, measurements are never defined at a instant of time.
+The second reason is that, upon close inspection, measurements are never defined at an instant of time.
 Therefore $x(t)$, the state at an instant of time, is really a shorthand for $x[(t - dt/2, t + dt/2)]$, the state defined over a small open interval $dt$ around $t$.
-We can talk about a single state within that interval only if all state variables do not change and the system is fully characterized by those state variables alone.
-Therefore writing $x(t)$ implicitly assumes two things: that the system is an equilibrium of the processes that happen at a faster scale of our measurements of said system; that further description (e.g.
-the internal dynamics of the constituents at a finer scale) is not needed to describe the process.
+We can talk about a single state within that interval only if no state variables change and the system is fully characterized by those state variables alone.
+Therefore writing $x(t)$ implicitly assumes two things: that the system is an equilibrium of the processes that happen at a faster scale than our measurements of said system; that further description (e.g.
+the internal dynamics of the constituents at a finer resolution) is not needed to describe the process.
 
@@
 
-All three insights points toward the same directly: the very notion of state seems to be intrinsically associated to the notion of equilibrium and state entropy.
+All three insights point toward the same direction: the very notion of state seems to be intrinsically associated to the notion of equilibrium and state entropy.
 This would suggest that all states for all systems are, in a sense, equilibria.
-But what is exactly this sense?
+But what exactly is this sense?
 What does it mean that a mechanical system, classical or quantum, is at equilibrium?
 
-Even in thermodynamics, equilibrium is usually an ill-defined concepts.
-It is typically associated the one coming from dynamical systems (i.e.
+Even in thermodynamics, equilibrium is usually an ill-defined concept.
+It is typically associated with the one coming from dynamical systems (i.e.
 the system no longer changes), but this doesn't quite work.
@@
 If we have a volume of gas in thermodynamic equilibrium, we would like to say it is in equilibrium for any observer, not just the one in its rest frame.
-A booster observer, then, would see an equilibria where the position is indeed changing.
-In presence of gravitational forces, this motion would not even be linear.
+A boosted observer, then, would see an equilibrium where the position is indeed changing.
+In the presence of gravitational forces, this motion would not even be linear.
 The notion of equilibrium as ``nothing is changing'', then, is not quite right and should be amended.
 But how?
-The previous discussion would hint toward system independence, but why system independence should be an equilibrium?
+The previous discussion would hint toward system independence, but why should system independence imply an equilibrium?
 
@@
 Therefore we need, at least conceptually, a process that takes that something and makes it independent from the rest, decouples it, makes it autonomous.
-Our ability to define a system implicitly assumes the existence of such ``decoupling'' process.
-Note that if the system is already independent, the decoupling process does nothing: in this sense independent systems are an equilibria of a decoupling process.
+Our ability to define a system implicitly assumes the existence of such a ``decoupling'' process.
+Note that if the system is already independent, the decoupling process does nothing: in this sense independent systems are an equilibrium of a decoupling process.
 
@@
 We cannot know how the system came to be in that state, we cannot know anything more about the subsystems, other systems or the environment than what the state of the system strictly tell us.
-All possible ways that the state can be realized must be realized in a possible evolution, or we could infer something more.
 When the system is independent, the states tell us the least possible information about everything else and the evolution of anything else.
@@
 an invariant, an equilibrium) of a group (i.e.
-a set of transformation, a set of processes).
+a set of transformations, a set of processes).
 
@@
 We can only study the state of objects that allow process entropy maximization.
-What remains is to answer one last fundamental question: is state entropy at all related to thermodynamic entropy?
+What remains is to answer one last fundamental question: is this notion of state entropy at all related to thermodynamic entropy?
 
@@
 The idea is that we want to describe the situation in which two systems come together, evolve as a composite for some time, and then proceed to evolve separately.
-In this case, at each juncture we have to reconcile the evolutions of both the whole and each of the part.
+In this case, at each juncture we have to reconcile the evolutions of both the whole and each of the parts.
 We need to have an independence equilibrium at both levels.
@@
 
-We say a system is in \textbf{thermodynamic equilibrium} if:
-\begin{itemize}
-	\item it is at equilibrium
-	\item all its parts are in thermodynamic equilibrium
-	\item if another system joined with a part of the system in thermodynamic equilibrium is in thermodynamic equilibrium, then the whole is in thermodynamic equilibrium.
-\end{itemize}
-We say two or more systems are in thermodynamic equilibrium if their composite is in thermodynamic equilibrium.
-A \textbf{thermodynamic process} is a process with equilibria where the final state is a thermodynamic equilibrium.
-We stress that we are not saying that two systems, in general, always allow for thermodynamic equilibria.
-However, we contend that these are the type of processes we can mix and match to create composite processes of composite systems.
-What we want to explore is what happens if we had such systems and process, at least to a first approximation.
-
-To get a feel for the definition, let us rederive the fact that thermodynamic equilibrium is transitive, also known as zeroth law.
-Suppose system $A$ is in thermodynamic equilibrium with system $B$ and system $B$ is in thermodynamic equilibrium with system $C$.
-Then the composite $B \times C$ is in thermodynamic equilibrium by definition.
-Since $A$ is in thermodynamic equilibrium with $B$, a part of $B \times C$, it is also in thermodynamic equilibrium with $B \times C$.
-Therefore $A \times B \times C$ is in thermodynamic equilibrium.
-But $A \times C$ is a part of $A \times B \times C$, so it is in thermodynamic equilibrium.
-Which means $A$ is in thermodynamic equilibrium with $C$.
-
-The characterization of thermodynamic equilibrium we give is manifestly a part-whole relationship.
-It extends the notion of independence equilibrium to a composite system so that the equilibrium of the whole and the equilibrium of its parts are essentially the same equilibrium.
-We find that clarifying this kind of part-whole relationships make for a good conceptual starting point not only for thermodynamics, but also for the other branches of physics.
-This should not come as a surprise as reductionism (i.e.
-explaining the whole through its parts) plays an important role in science.
-
-Another important question for a part-whole relationships is: how are the state variables of the whole related to the state variables of the parts?
+If our goal is to characterize part-whole relationships, a crucial question is: how are the state variables of the whole related to the state variables of the parts?
 We define a \textbf{thermodynamic system} one that follows these rules:
@@
 \end{align}
-which are the usual thermodynamic relationships, except that, for us, the state entropy $S$ dimensionless..
+which are the usual thermodynamic relationships, except that, for us, the state entropy $S$ is dimensionless.
+
+As we need to characterize states of both whole and parts, we need to characterize the part-whole relationship of the equilibrium.
+We say that a thermodynamic system $A$ in \textbf{thermodynamic equilibrium} has the following properties:
+\begin{itemize}
+	\item the system $A$ is in an independence equilbrium
+	\item all parts of $A$ are in thermodynamic equilibrium
+	\item if another composite system $B$ is in thermodynamic equilibrium and $A$ and $B$ share a part, then the composite of $A$ and $B$ is in thermodynamic equilibrium.
+\end{itemize}
+We say two or more systems are in thermodynamic equilibrium with each other if their composite is in thermodynamic equilibrium.\footnote{To understand the mathematical structure a bit better, the set of all systems $S$ is a Boolean algebra.
+Element $0$ represents ``nothing'' while element $1$ represents ``everything''.
+The join $\OR$ operation is system composition while the meet $\AND$ returns the smallest system contained by the two.
+Therefore $A \OR 1 = 1$: any system composed with everything gives us everything.
+As a lattice, the partial order $\leq$ is system inclusion.
+A thermodynamic equilibrium essentially identifies a topmost system, with all the subsystem and only the subsystems as part of the same equilibrium.
+That is, for every thermodynamic equilibrium there exists a system $A$ such that every $B \leq A$ is in the same equilibrium while every $C \nleq A$ is not in the same equilibrium.} A \textbf{thermodynamic process} is a process with equilibria where the final state is a thermodynamic equilibrium.
+We stress that we are not saying that two systems, in general, always allow for thermodynamic equilibria.
+However, we contend that these are the type of processes we can mix and match to create composite processes of composite systems.
+What we want to explore is what happens if we had such systems and processes, at least to a first approximation.
+
+To get a feel for the definition, let us rederive the fact that thermodynamic equilibrium is transitive, also known as the zeroth law.
+Suppose system $A$ is in thermodynamic equilibrium with system $B$ and system $B$ is in thermodynamic equilibrium with system $C$.
+Then the composites $A \times B$ and $B \times C$ are respectively in thermodynamic equilibrium.
+Since they share a part, $B$, this means that $A \times B \times C$ is in thermodynamic equilibrium.
+But $A \times C$ is a part of $A \times B \times C$, so it is in thermodynamic equilibrium.
+Which means $A$ is in thermodynamic equilibrium with $C$.
+
+The characterization of thermodynamic equilibrium we give is manifestly a part-whole relationship.
+It extends the notion of independence equilibrium to a composite system so that the equilibrium of the whole and the equilibrium of its parts are essentially the same equilibrium.
+We find that clarifying this kind of part-whole relationships makes for a good conceptual starting point not only for thermodynamics, but also for the other branches of physics.
+This should not come as a surprise as reductionism (i.e.
+explaining the whole through its parts) plays an important role in science.
 
@@
 
-A natural question, based on the previous definition, is: to what extend thermodynamic processes can change the energy and the entropy of each system?
+A natural question, based on the previous definitions, is: to what extent can thermodynamic processes change the energy and the entropy of each system?
 Intuitively, there are two ``orthogonal'' directions for a change of energy: constant entropy and maximal entropy change.
@@
 The energy of the mechanical system is effectively \emph{stored} in one of the other state variables and can be later retrieved.
-We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U$ as the energy gained by the system during a thermodynamic process.\footnote{Note that $\frac{\partial S_M}{\partial U} = 0 = \beta_M = \frac{1}{k_B T_M}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite.
+We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U$ as the energy gained by the system during a thermodynamic process.\footnote{Note that $\frac{\partial S}{\partial U} = 0 = \beta = \frac{1}{k_B T}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite.
 While this ``temperature'' has little to do with thermometers, the fact that we can always transform all work into heat but not vice-versa is just a specific case of being able to move energy from higher to lower temperature.}
@@
 Consider a system composed of a generic system $A$, a thermal reservoir $R$ and a mechanical system $M$.
-In this setting, work represents the energy exchanged at constant entropy while heat represents energy exchanged associated with entropy change.
+In this setting, work represents the energy exchanged at constant entropy while heat represents exchanged energy associated with entropy change.
 
@@
 
-We now look at the change in entropy, which during a thermodynamic process can never decrease. Keeping in mind $M$ is mechanical, we have:
+We now look at the change in entropy, which during a thermodynamic process can never decrease given that thermodynamic processes are deterministic. Keeping in mind $M$ is mechanical, we have:
 \begin{equation}
@@
 &\Delta S_A \geq - \Delta S_R \\
-&\Delta S_A \geq \int \beta_R dQ \\
-&\Delta S_A \geq \int \frac{dQ}{k_B T_R}.
+&\Delta S_A \geq \beta_R \Delta Q \\
+&\Delta S_A \geq \frac{\Delta Q}{k_B T_R}.
 \end{aligned}
@@
 If the ground state is non-degenerate, such as a perfect crystal, its entropy is equal to zero.
-We can easily recover it in this form.
+We can more easily recover it in this form.
 
-Recall that state entropy counts the logarithm of the possible evolutions of the system at equilibrium.
-The lowest possible count $\mu$ of evolution is one, which is achieved if the equilibrium can be reached only by one starting condition.
+Recall that state entropy is the logarithm of the count of the possible evolutions of the system at equilibrium.
+The lowest possible count $\mu$ of evolutions is one, which is achieved if the equilibrium can be reached only by one starting condition.
 In this case, $S=\log(\mu)=0$.
 This tells us that state entropy is non-negative and therefore is bounded from below.
-We now have to show that all thermodynamic systems must have a lowest energy state and that it must be a local minima for entropy.
+We now have to show that all thermodynamic systems must have a lowest energy state and that it must be a local minimum for entropy.
 
@@
 Now we show that either all thermodynamic systems must have a lowest energy state or they must have a highest energy state.
-Suppose, in fact, we had two system, $A$ has no maximum energy while $B$ has no minimum energy.
+Suppose, in fact, we had two systems: $A$ has no maximum energy while $B$ has no minimum energy.
 Since $S_A(U_A)$ is concave, it must be a monotonically increasing function of $U_A$.
@@
 Conversely, $S_B(U_B)$ must be monotonically decreasing, which means $\beta_B$ is always negative.
-If we put the two system in contact, the energy of $B$ will keep decreasing and the energy of $A$ will keep increasing and no equilibrium will ever be reached.
+If we put the two systems in contact, the energy of $B$ will keep decreasing and the energy of $A$ will keep increasing and no equilibrium will ever be reached.
 Therefore all thermodynamic systems either have a minimum or a maximum energy.
@@
 To finish, we note that a perfect crystalline structure fits the case where the system has only one possible evolution, so it is a zero entropy state.
-If there is degeneracy, there are different ways the equilibrium can be realized, there are different evolutions that will correspond to the same equilibrium which means the entropy will be greater than zero.
+If there is degeneracy, there are different ways the equilibrium can be realized, instead, there are different evolutions that will correspond to the same equilibrium which means the entropy will be greater than zero.
 This recovers the third law.
 
-The laws of thermodynamics have been recovered, and we have done so only by characterizing how equilibria and state variables behave under system combination.
+The laws of thermodynamics have been recovered, and we have done so only by characterizing how equilibria and state variables behave under system composition.
 Apart from that, we have made no assumptions on what type of system, the type of forces acting on the system, the internal composition of the system or the internal dynamics of the system.
 No reference to statistical mechanics or quantum mechanics is needed: the theory stands on its own.
-This fits the original spirit of the thermodynamics, which was to find a set of laws that would be independent on the ultimate dynamics of the constituents.
-Additionally, the concepts we used are directly related to the physical explanation, which is not recovered after pages and pages of calculation.
+This fits the original spirit of thermodynamics, which was to find a set of laws that would be independent fn the ultimate dynamics of the constituents.
+Additionally, the concepts we used are directly related to the physical explanation, and the laws are recovered immediately, not after pages and pages of calculations.
 The combination of these features makes this approach directly useful, for example, in an introductory physics course: a coherent self-contained explanation can be given in terms of simple definitions.
@@
 
-While recovering thermodynamics from straight forward physical definitions has its own merits, we believe the approach can be used to offer gave additional insights new insights on classical mechanics, quantum mechanics and statistical mechanics.
+While recovering thermodynamics from straightforward physical definitions has its own merits, we believe the approach can be used to offer new insights into classical mechanics, quantum mechanics and statistical mechanics.
 So we want to conclude with a few remarks on the subject that give a feel for how we intend to extend the work in the fully developed framework.
@@
 Is this warranted and useful?
-Do all mechanical system match our requirement to have constant entropy?
-If one looks closely, these facts are implicit assumed.
-For quantum mechanics, in fact, the Von Neumann entropy of every pure state is zero.
+Do all mechanical systems match our requirement to have constant entropy?
+If one looks closely, these facts are already implicitly assumed.
+For quantum mechanics, in fact, the von Neumann entropy of every pure state is zero.
 In classical mechanics, a uniform measure is used to count the accessible states; additionally the Gibbs entropy is invariant under translations in phase space.
-It is also embedded directly in the corresponding mechanics.
-In both theories, given two states, we can always (at least mathematically) find a deterministic and reversible evolution (i.e.
-a Hamiltonian evolution or a unitary evolution) that connects the two.
-So the idea that mechanical systems are associated with a constant entropy, though seldom explicit, can be already deduced from the currently established theories.
+This means that all states must contribute the same entropy.
+Additionally, in both theories, given two states, we can always (at least mathematically) find a deterministic and reversible process (i.e.
+a Hamiltonian evolution or a unitary evolution) that starts with the first and ends in the second.
+So the idea that mechanical systems are associated with a constant entropy, though seldom explicit, can already be deduced from the currently established theories.
 
-Another natural question is: how does the process entropy relates to the fundamental postulate of statistical mechanics, that entropy is the logarithm of the count of states?
+Another natural question is: how does the process entropy relate to the fundamental postulate of statistical mechanics, that entropy is the logarithm of the count of states?
 Suppose we give the description of a system at a given time (e.g.
@@
 Also, this is the sum (or integral) of the process entropy associated with each state not ruled out by that description: every evolution will have to go through one of those states at that time.
-Now suppose the system is, at that time, independent from the other.
+Now suppose the system is, at that time, independent from all others.
 Then the process entropy for each state is the state entropy, so the process entropy for that description is the sum (or integral) of the state entropy.
@@
 We could have an independent mechanical system under a deterministic and reversible process.
-Then counting evolution is the same as counting states, up to a constant, at any moment in time.
-In another scenario, the process is not deterministic and reversible but is isoentropic, meaning that the process entropy is constant in time.
-If we have an independent mechanical system, its evolution will necessarily be isoentropic since, whatever state we evolve to, the state entropy will always be the same.
-Now suppose the set of states compatible with our description mix with each other and only with each other.
+Then counting evolutions is the same as counting states, up to a constant, at any moment in time.
+In another scenario, the process is not deterministic and reversible but is isentropic, meaning that the process entropy is constant in time.
+If we have an independent mechanical system, its evolution will necessarily be isentropic since, whatever state we evolve to, the state entropy will always be the same.
+Now suppose the states compatible with our description mix with each other and only with each other.
 Each evolution will change state in time, but it will never go outside of the set, so the description works both as a constraint and a description of the dynamic equilibrium.
@@
 The process entropy at the initial state will coincide with the count of states, but not at later times.
-As the evolution concentrate around the equilibrium, the number of states become fewer so the number of evolutions per state increases.
+As the evolutions concentrate around the equilibrium, the number of states becomes fewer so the number of evolutions per state increases.
 The entropy of the equilibrium, assuming that it is ever reached, is essentially the size of the basin of attraction at the initial time.
@@
 Conversely, note that the areas in phase space shrink as the available states become fewer and fewer, which would mean entropy as the log of the state count decreases under an irreversible process.
-It was actually this paradox that pushed into considering evolutions as more fundamental than states and, retrospectively, it is the obvious choice.
+It was actually this paradox that pushed us into considering evolutions as more fundamental than states and, retrospectively, it is the obvious choice.
 
-This example also show how critical is the notion of independence to define state entropy.
+This example also shows how critical the notion of independence is to define state entropy.
 The evolution of a damped harmonic oscillator is not independent from the evolution of the systems that are absorbing the dissipated energy, which are in fact correlated: the higher the energy of the oscillator, the higher the energy lost.
@@
 In this paper we gave a conceptual overview for a new approach to recover thermodynamics based on the simple idea of counting evolutions, the possible ways a system can evolve under a process.
-We saw that the logarithm of evolution count under a deterministic process has all the properties we associate with entropy and can serve as a foundations for thermodynamics but also as a common foundations for the other physical theories.
+We saw that the logarithm of evolution count under a deterministic process has all the properties we associate with entropy and can serve as a foundation for thermodynamics but may also be further developed into a common foundation for the other physical theories.
 
 Our aim is to fully develop a formal framework based on these ideas, which will give us a general theory for states and processes.
-Conceptually, we start with all that can be observed at all moments of time; organize them into systems/degrees of freedom by grouping those that ``go together'' in the sense that they can be studied, at least under some circumstances, independently from the rest; define state spaces as ``templates'' of those descriptions that can be instantiated over time.
-This will force us to spell out the hidden assumptions one always have when defining systems, states and processes.
-The goal is to then recover the different physical theories with additional assumptions on the system or on the processes being studied.
+Conceptually, we start with all that can be observed at all moments of time; organize the possible descriptions into systems/degrees of freedom by grouping those that ``go together'' in the sense that they can be studied, at least under some circumstances, independently from the rest; define state spaces as ``templates'' of those descriptions that can be instantiated over time.
+This will force us to spell out the hidden assumptions one always has when defining systems, states and processes.
+The goal is to then recover the different physical theories with additional assumptions on the system or on the processes being studied, as we have done here for thermodynamics.
 
@@
 topologies, $\sigma$-algebras).
-This need to be extended to allow us to count possibilities in general and evolutions in particular.
-Technically, we need a unified tool that generalizes notions of measure theory, geometry, probability and information theory as these cover specific aspect that do not work in the general case.
+This needs to be extended to allow us to count possibilities in general and evolutions in particular.
+Technically, we need a unified tool that generalizes notions of measure theory, geometry, probability and information theory as these cover specific aspects that do not work in the general case.
 
 It is important to us that these additional efforts, while making the mathematics more precise, will not obscure the physical motivation and intuitive understanding of the framework.
-Mathematical consistency is a tool We believe there is a way
-
 These additional efforts will not change the nature of the insights presented.

### 2021-01-15 Articles/2019-ShannonEntropy/Submissions/EJP/response.tex
Incorporating feedback from Julian

@@
 \end{quote}
-Again, it is difficult to understand whether, hidden under all the ad hominem, there is an actual point. The word `perplexing' is used in the following paragraph:
+Again, it is difficult to understand whether, hidden under the personal attacks, there is an actual point. The word `perplexing' was used in the following paragraph:
 
@@
 This would leave us very perplexed...
+
+We modified the first sentence to read:
+
+\emph{This may seem particularly perplexing given that we want to use this quantity in a physical setting.}
 
@@
 They are called \textbf{defining constants} because they define the unit system, which is a convention.
-Before engaging in comments that imply we are incompetent, the referee may want to read the relevant literature, which are now included in the paper.
+Before engaging in comments that imply we are incompetent, the referee may want to read the relevant literature, which is now included in the paper.
 
@@
 \emph{In the latest revision of the International Metric System of Units SI, the Boltzmann constant is one of the defining constants.
-This means its value is assigned, rather then measured, in a way that properly defines Kelvin, the unit of temperature.
+This means its value is assigned, rather than measured, in a way that properly defines Kelvin, the unit of temperature.
 The universal constants are essentially used as fixed starting points to define our measurement scales.}

### 2021-01-15 Articles/2019-ShannonEntropy/Submissions/EJP/EJP-ShannonEntropyIsVariability.tex
Incorporating feedback from Julian

@@
 
-This should be particularly perplexing if we want to use this quantity in a physical setting.
+This may seem particularly perplexing given that we want to use this quantity in a physical setting.
 If we want $\rho(x)$ to represent the state of a macroscopic system and $H(\rho(x))$ to represent entropy, a state variable, how can $H$ change value if we merely express $\rho$ over different coordinates?
@@
 The Boltzmann constant $k_B$ should not distract us: its role is simply to allow us to measure temperature in an appropriate unit.\footnote{In the latest revision of the International Metric System of Units SI,\cite{chyla2011evolution} the Boltzmann constant is one of the defining constants.
-This means its value is assigned, rather then measured, in a way that properly defines Kelvin, the unit of temperature.
+This means its value is assigned, rather than measured, in a way that properly defines Kelvin, the unit of temperature.
 The universal constants are essentially used as fixed starting points to define our measurement scales.

### 2021-01-18 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Edited Section 4 and rewrote 5.3 (Curiel)
Curiel section still has some issues. And still some questions for section 4.2 stuff (and at least one thing to double check in 4.1, indicated with XXX, p. 16).

@@
 
+% % To do: slip in a citation to Dorr/Hawthorne 2014, to note the independent metaphysical challenges that the notion of naturalness faces, and likewise for a metaphysical notion of fundamentality.
+
 % North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory.
@@
 For instance, we can think of the canonical position and momentum observables as ``epistemically fundamental" because they suffice to completely specify a classical state, fixing its values for all other observables.
-\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{density}, discusses another sense in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied.
+\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{fundamental}, discusses further senses in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied.
 Compared with North's \parencites*[]{North} account of structure, our definition has two distinct advantages.
@@
 
-Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.
-While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus structure.
+Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus structure.
 Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration.
@@
 
-Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as formal device for encoding physical information or even merely as a calculational tool.} Whether Hamiltonian phase space structure is metaphysically fundamental does not matter for the questions we seek to answer.
+Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as a formal device for encoding physical information or as a calculational tool.} Whether Hamiltonian phase space structure is metaphysically fundamental does not matter for the questions we seek to answer.
 Regardless of its metaphysical status, Hamiltonian mechanics is physically more fundamental, and we will argue for its epistemic and conceptual primacy as well.
-[To do: only keep preceding remark if somewhere later in the essay, we elucidate these epistemic and conceptual dimensions.] Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references].
+Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references].
 Each of these frameworks has different pragmatic costs and benefits.
@@
 There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions.
-Furthermore, each of these formalisms references some ``unphysical" coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.
+Furthermore, each of these formalisms references some ``unphysical" coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.\footnote{Tulczyjew's reformulation of Hamiltonian mechanics supplies a further reason for viewing symplectic geometry---and hence Hamiltonian mechanics---as conceptually fundamental.
+According to Teh and Tsementzis, this reformulation shows that  ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh}.}
 %The mathematical structure that you ``need" might not be unique.
@@
 North also introduces a technical notion of comparative structure, based on the symmetry group of statespace.
-According to this criterion, statespace A has more structure than statespace B if the symmetry group of A is a proper subgroup of B's symmetry group \parencites[87-88]{North}.\footnote{This statement follows the charitable reconstruction of North's criterion given by \textcites[]{Swanson}.
+According to this criterion, statespace A has more structure than statespace B if the symmetry group of A is a proper subgroup of B's symmetry group \parencites[87-88]{North}.\footnote{This statement follows the most charitable reconstruction of North's criterion given by \textcites[]{Swanson}.
 North sometimes speaks in terms of comparing the dimensions of symmetry groups, which they note leads to trivial counterexamples.} A smaller symmetry group entails that it is in some sense ``harder" to preserve the space's structure, indicating that it has more mathematical structure to be preserved.
@@
 We integrate yet again and find:
-\begin{equation}
+\begin{equation} \label{massive_Hamiltonian}
 H = \frac{1}{2m} (p_j - A_j) g^{ij}(p_j - A_j) + V \\
@@
 
-To understand what makes densities special in the context of Hamiltonian mechanics, it is helpful to consider densities in general.
-By a `density' we mean the intuitive physical notion of a quantity divided by the region it occupies, like mass per volume or charge per area.
+To understand what makes densities special in the context of Hamiltonian mechanics, we first consider densities in general.
+By a `density' we mean the intuitive physical notion of a quantity divided by the region it occupies, such as mass per volume or charge per area.
 Over a continuum, this is an integrable function $\rho(x)$.
@@
 In differential geometry, a different mathematical object is called the `density,' namely a volume form (i.e.
-an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation.\footnote{TODO: The differential density can be understood as a way to generalize volume forms in the case of a non-orientable manifold.
-This case is not interesting to us, therefore we will be glossing over issues of orientation and subtle differences between volume forms and differential densities.
-We believe this simplifies the exposition removing details that have no real consequence for the physical understanding.} Our physically significant density $\rho$ is actually the component of this volume form $\nu = \rho(x^i)  \epsilon dx^1 \wedge dx^2 \wedge ...
-\wedge dx^n$ where $\epsilon$ is $1$ or $-1$ depending on the orientation..
+an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation.\footnote{The differential density generalizes volume forms in the case of a non-orientable manifold.
+However, we focus only on cases where the manifold is orientable.
+To simplify the exposition, we will elide subtle differences between volume forms and differential densities.} Our physically significant density $\rho$ is actually the component of this volume form $\nu = \rho(x^i)  \epsilon dx^1 \wedge dx^2 \wedge ...
+\wedge dx^n$ where $\epsilon$ is $1$ or $-1$ depending on the orientation.
 Note that $\nu$ combines the density and the volume into a coordinate-free object.
-Considering the total mass in the region $U$, we have $m(U) = \int \nu$ so essentially $\nu = dm$ which is the infinitesimal mass contribution we are integrating.
+Considering the total mass in the region $U$, we have $m(U) = \int \nu$ so essentially $\nu = dm$, i.e.
+the infinitesimal mass contribution we are integrating.
 But this means that the value returned by $\nu$ depends not only on how much stuff there is at a given point, but also on the size of the volume identified by the $n$ vectors given to the form.
@@
 At first glance, this expression misleadingly suggests that the density $\hat{\rho}$ is coordinate-independent: it is a scalar quantity with the correct dimensions of mass over volume.
-[XXX] Nevertheless, $\hat{\rho}$ remains coordinate-dependent because it is a function of $|g_{ab}|$, which depends on the choice of coordinates and is, in general, a function of the point.
-However, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
+[XXX to do: ask GC if the following explanation is accurate] Nevertheless, $\hat{\rho}$ remains coordinate-dependent because specifying a numerical value for $\hat{\rho}$ requires choosing a particular set of coordinates.
+Fortunately, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
 Since the square root of the determinant $|g_{ab}|$ is the Jacobian of the transformation to those coordinates, $|g_{ab}|$ will be constant only when this Jacobian equals one.
 For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$; in polar coordinates $|g_{ab}| = r^2$ and hence the volume form $dV = r dr \wedge d\theta$.
-This means that $\hat{\rho}$ at best expresses the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise.\footnote{This is similar to what happens with the concept of proper time: it is defined in the comoving frame, so it is coordinate dependent.
-Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate invariant.} Again, the upshot here is that we cannot avoid using a coordinate-dependent density.
+This means that $\hat{\rho}$ at best expresses the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise.\footnote{This is similar to what happens with the concept of proper time: it is defined in the co-moving frame, so it is coordinate dependent.
+Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate invariant.
+This coordinate-invariance derives from the fact that we have conventionally agreed to define proper time in terms of the object's co-moving frame, rather than in terms of a different frame.} By conventionally agreeing to work in a coordinate system where $|g_{ab}| = 1$, we can make the density $\hat{\rho}$ coordinate-invariant.
+But, the property of point-wise comparison remains coordinate-dependent: it requires working in a particular coordinate system.
 Outside the context of Hamiltonian mechanics, the best we can do is obtain the property of point-wise comparison in a special coordinate system.
+% General idea from one-11 meeting: in the general case, in order to make the density a coordinate-invariant [so that we all can agree on its value], we have to make a global choice of coordinate system.
+We need to choose coordinate such that the determinant of the metric tensor is equal to 1.
+So we are seemingly saying that there are coordinate systems that are special, namely those where we can do point-wise comparison.
+So this poses a problem for the view that nature does not privilege one coordinate system over another.
+Possible response from the metaphysicians: perhaps the real fundamental physical object is a quotiented-out version of the density.
+
+% From before 1-11 : Again, the upshot here is that we cannot avoid using a coordinate-dependent density. 
 %We can also see this if we look at the units of $\hat{\rho}$, which are in terms of inverse volume as expressed in the privileged system.
 From 1-4 meeting: I thought that $\hat{\rho}$ has units of mass over volume.
+% Original bad explanation (see 1-11-21 meeting, around 11:13 am): Nevertheless, $\hat{\rho}$ remains coordinate-dependent because it is a function of $|g_{ab}|$, which depends on the choice of coordinates and is, in general, a function of the point.
 
@@
 \wedge dv^n = |g_{ab}| \bigwedge dx^i \bigwedge dv^j$.
-The determinant of the metric tensor is therefore directly the component of the volume form of the tangent bundle.
+Hence, the determinant of the metric tensor is equals the component of the volume form of the tangent bundle.
 The relationship to the density as expressed in phase space is therefore the following:
@@
 The left side will remain invariant.
-The two terms on the right side will change in the opposite way so that their product is invariant.
-Focusing on densities, then, brings about a physically-motivated way to understand the relationship between the cotangent and the tangent bundle, where the metric tensor plays a role in defining the geometry of the second in a similar way that the symplectic form defines the geometry of the first: it defines the count of states so that densities can be properly tracked.
-
-Overall, these considerations show that in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
+The two terms $|g_{ab}|$ and  $\rho(x^i, v^j)$ on the right side will change in the opposite way so that their product is invariant.
+Focusing on densities, then, brings about a physically-motivated way to understand the relationship between the cotangent and the tangent bundle, where the metric tensor plays a role in defining the geometry of the latter similar to how the symplectic form defines the geometry of the former: it defines the count of states so that densities can be properly tracked.
+
+These relationships also allow us to better understand the hyper-regular case, and in particular the case of a massive particle.
+Note that, in general, the relationship between densities over the cotangent and tangent bundles is, as we saw:
+\begin{equation} \label{densities_relationship}
+\rho(q^i, p_j) = \left|\frac{\partial v^j}{\partial p_k}\right| \rho(x^i, v^j)  = \left|\frac{\partial H}{\partial p_k \partial p_j}\right| \rho(x^i, v^j).
+\end{equation}
+In the case of massive particles, where the Hamiltonian (and the Lagrangian) are quadratic, the Hamiltonian of Equation~\ref{massive_Hamiltonian} entails that $\frac{\partial H}{\partial p_k \partial p_j} = m g^{kj}$.
+This is the case where the densities in the tangent bundle can be properly understood in terms of Riemmanian geometry.
+For the more general case, we note that
+\begin{equation}
+\begin{aligned}
+\delta^i_j = \frac{\partial v^i}{\partial v_j} = \frac{\partial v^i}{\partial p_k}\frac{\partial p_k}{\partial v_j} &= \frac{\partial}{\partial p_k}\left(\frac{\partial H}{\partial p_i}\right) \frac{\partial }{\partial v_j} \left( \frac{\partial L}{\partial v_k} \right) \\
+\Rightarrow \frac{\partial L}{\partial v^k \partial v^j} &= \left[ \frac{\partial H}{\partial p_k \partial p_j} \right]^{-1}
+\end{aligned}
+\end{equation}
+which means that the Hessian of the Hamiltonian is the inverse of the Hessian of the Lagrangian, and so are their determinants.
+Recall that the hyper-regular case corresponds to the case where the determinant of the Hessian of the Lagrangian is non-zero.
+Hence, Equation~\ref{densities_relationship} tells us that in this case, we can express the density of states over position and velocity.
+
+To summarize the main theme from this section: in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
 In this context, we need to know which coordinate system we are in before making these comparisons.
@@
 Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
-% Josh: why is the determinant of the metric tensor equal to the volume of the tangent bundle in this case? [Discussed on 1-4 meeting, 11:56 a.m.]. Perhaps study discussion of this in the long paper.
-
-[XXX] These relationships also allow us now to better understand what the hyperregular case, and the massive particle in particular, correspond to.
-Note that, in general, the relationship between densities over the tangent and cotangent bundles is, as we saw:
-\begin{equation}
-\rho(q^i, p_j) = \left|\frac{\partial v^j}{\partial p_k}\right| \rho(x^i, v^j)  = \left|\frac{\partial H}{\partial p_k \partial p_j}\right| \rho(x^i, v^j).
-\end{equation}
-Note that $\frac{\partial H}{\partial p_k \partial p_j} = m g^{kj}$ exactly in the case of massive particles, where the Hamiltonian (and the Lagrangian) are quadratic.
-This is the case where the densities in the tangent bundle can be properly understood in terms of Riemmanian geometry.
-For the more general case, we note:
-\begin{equation}
-\begin{aligned}
-\delta^i_j = \frac{\partial v^i}{\partial v_j} = \frac{\partial v^i}{\partial p_k}\frac{\partial p_k}{\partial v_j} &= \frac{\partial}{\partial p_k}\left(\frac{\partial H}{\partial p_i}\right) \frac{\partial }{\partial v_j} \left( \frac{\partial L}{\partial v_k} \right) \\
-\frac{\partial L}{\partial v^k \partial v^j} &= \left[ \frac{\partial H}{\partial p_k \partial p_j} \right]^{-1}
-\end{aligned}
-\end{equation}
-which means the Hessian of the Hamiltonian is the inverse of the Hessian of the Lagrangian, and so are their determinants.
-The hyperregular case corresponds to the case where the determinant of the Hessian of the Lagrangian is non-zero, which means we can express the densities of states over position and velocity.
 
@@
 
-We now explain why the conjugate momentum $p$ is gauge-dependent.
-[XXX: first we want to answer: what is a gauge transformation?
-What does it represent physically?] Recall that motion equivalence entails the following: knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
+We now consider transformations that leave the density, position, and velocity invariant.
+We will see that we can interpret these transformations as a set of gauge transformations on the conjugate momentum $p $.
+This entails that---unlike both the canonical position state variable $q$ and the kinetic momentum---the conjugate momentum is a gauge-dependent quantity.
+Recall that motion equivalence entails the following: knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
 In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$.
 Hence, any transformation that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant.
-We want such transformations to be canonical in order to preserve the invariant density.
-This means that the new state variables $(q, \hat{p})$ will respect the usual Poisson brackets:
+We will further assume that the transformations leave the density invariant, which entails that the new variables must be canonical.
+Hence, the new state variables $(q, \hat{p})$ must respect the usual Poisson brackets:
 \begin{equation}
@@
 	\{q^i, \hat{p}_j\} &= \frac{\partial q^i}{\partial q^k} \frac{\partial \hat{p}_j}{\partial p_k} - \frac{\partial q^i}{\partial p_k} \frac{\partial \hat{p}_j}{\partial q^k} \\
-	&= \delta^i_k \frac{\partial \hat{p}_j}{\partial p_k} - 0 \frac{\partial \hat{p}_j}{\partial q^k} = \frac{\partial \hat{p}_j}{\partial p_i} = \delta^i_j
+	&= \delta^i_k \frac{\partial \hat{p}_j}{\partial p_k} - 0 \frac{\partial \hat{p}_j}{\partial q^k} \Rightarrow \frac{\partial \hat{p}_j}{\partial p_i} = \delta^i_j
 \end{aligned}
 \end{equation}
-Integrating we have:
+Integrating this last equation yields
 \begin{equation}
@@
 Since $A_j$ is curl free, we can express it as the gradient of a function $f$.
-Therefore changes of state variable of the form $\hat{p}_j = p_j + \partial_j f$ do not change either the equations of motion or the trajectories.
-We recognize these as a family of gauge transformations that change the conjugate momentum variable $p$ while leaving $(q,\dot{q}) $ invariant.
-[XXX: make it clear that we found gauge transformations with our different definition] The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
+Therefore, changes of state variable of the form $\hat{p}_j = p_j + \partial_j f$ do not change either the equations of motion or the trajectories.
+From the usual notion of a gauge transformation, we recognize these as a family of gauge transformations that change the conjugate momentum variable $p$ while leaving $(q,\dot{q}) $ invariant.
+We can thereby physically interpret this set of gauge transformations as the transformations that leave the density and $(q,\dot{q}) $ invariant.
+
+The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
 Additionally, it shows that not all state variables are on equal physical footing.
 Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.
-Indeed, as we now argue, some coordinate choices are more significant than others.
-[XXX: coordinate or state variables?]
+Indeed, as we now argue, some state variable choices are more physically significant than others.
 % Whereas symplectic geometry creates a mathematical, syntactic symmetry between the $q$ and $p$ variables, this is not a physically significant symmetry from the standpoint of measurement.
@@
 For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
-Yet, some choices of the coordinates $q$ are not appropriately connected with physical measurement, making them less physically meaningful [to do: provide a clear example of this].
-Crucially, the coordinates $q$ define the units of physical measurement, allowing us to assign numbers to physical quantities.
+Yet, some choices of the coordinates $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [to do: provide a clear example of this].
+The phase space density remains invariant upon changing the coordinates $q$, inducing a state variable transformation on the conjugate momenta $p$ (which compensate by transforming covariantly).
+It is in this sense that the coordinates $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.
 These units are physically significant despite not being coordinate-invariant.
@@
 [Not sure if/where to include the following point (possibly with the discussion of the coordinate-dependence of the density?): As the derivation in Section~\ref{deterministic} shows, a necessary condition for defining the symplectic structure is that we can express the density in a particular choice of units.
-If we did not have the physical ability to define the density in some units we would not be able to define this structure.
-]
-
-[To do: expand on the following point: ] This is why statistical mechanics works best on phase space (we have a uniform measure over q/p).
+If we did not have the physical ability to define the density in some units we would not be able to define this structure.]
+
+[To do: expand on the following point:] This is why statistical mechanics works best on phase space (we have a uniform measure over q and p).
 Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
@@
 They are coordinate invariant.
-Also the Shannon/Gibbs entropy are coordinate invariant.
+Also the Shannon and Gibbs entropies are coordinate invariant.
 
@@
 Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition.
-We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating the generalization, see \textcites[110]{Arnold1989}.
+We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating the generalization, see \textcites[110]{Arnold}.
 [To do: verify that Arnold can legitimately be read as generalizing the argument below.]}
@@
 
-(XXX: Note how the geometry is better understood in extended phase space, requiring Hamiltonian picture.
-Note that this applies even when $\theta S$ cannot be expressed as function of $q$ and $\dot q$.
+(XXX to do: Note how the geometry is better understood in extended phase space, requiring the Hamiltonian picture.
+Note that this applies even when $\theta \cdot S$ cannot be expressed as a function of $q$ and $\dot q$.
 )
@@
 
-North assumes that formulating a Lagrangian theory requires specifying a Riemannian metric; according to Curiel, this is neither necessary nor sufficient.
+North assumes that formulating a Lagrangian theory requires specifying a Riemannian metric [but this is not true: see the discussion of her appendix A above]; according to Curiel, this is neither necessary nor sufficient.
 Furthermore, the physical significance of a Riemannian metric in the context of a Lagrangian theory is unclear: two different Lagrangians can have the same Lagrangian vector field and yet induce inequivalent Riemannian metrics
@@
 Curiel frequently appeals to naturalness and a distinction between natural vs.
-ad hoc/artificial constructions [he defines `naturalness', specifically a `natural coordinate system', in terms of a particular kind of coordinate transformation.
+ad hoc/artificial constructions [he defines `naturalness', specifically a `natural coordinate system', in terms of a particular kind of coordinate transformation, using the notion of the almost tangent $J $ structure.
 See his equation 4.3, page 22 of the PDF].
@@
 
-\subsection{Response to Curiel's criticisms of Hamiltonian mechanics}
+\subsection{Curiel's privileging of Lagrangian mechanics}
 \label{Curiel}
 
+% Helpful summary of Curiel's position: ``Hamiltonian mechanics does not respect the kinematical constraints intrinsic to abstract classical systems" while ``Lagrangian mechanics always respect[s] the kinematical constraints of abstract classical systems" \parencites*[304]{Curiel}.
+
 Curiel \parencites*[]{Curiel} argues that Hamiltonian mechanics is less physically significant than Lagrangian mechanics.
-He attempts to show that the Lagrangian framework represents classical mechanical systems in a \textit{natural} matter whereas the Hamiltonian formulation involves an artificial construction \parencites*[270]{Curiel}.
-His argument hinges on the claim that Hamiltonian mechanics requires an ad hoc condition that is imposed by hand, whereas Lagrangian mechanics follows from physically plausible principles.
-[Set this up properly: Curiel's definition of abstract classical system; Curiel's privileging of certain kinematical constraints, his equations 4.3; Hamiltonian form of equation 6.2; see below].
+He attempts to show that the Lagrangian framework represents classical mechanical systems in a \textit{natural} manner whereas the Hamiltonian formulation involves an artificial construction \parencites*[270]{Curiel}.\footnote{Curiel defines `natural' in terms of natural coordinate systems, defined using the almost-tangent structure of the tangent bundle \parencites*[290-291]{Curiel}.
+He interprets this as capturing the fundamental kinematical constraints relating position and velocity in classical mechanics.} Central to Curiel's argument is the claim that Lagrangian mechanics naturally encodes the kinematical structure of classical systems whereas Hamiltonian mechanics requires imposing an \textit{ad hoc} condition.
+We will now argue that the condition Curiel claims is `ad hoc' actually has a natural physical interpretation.
+Additionally, we note a disagreement about which formal classical systems are physically significant.
+
+Similar to both us and North, one of Curiel's main goals is to identify the necessary and sufficient mathematical structures for formulating classical mechanics.
+Curiel argues that ``the kinematically-possible evolutions" of a classical system correspond exactly to ``the possible solutions to the Euler--Lagrange equation posed on the tangent bundle" of the system's configuration space \parencites*[292]{Curiel}.
+Hence, he sets out to establish the ``geometrical structure necessary (and manifestly sufficient) for the formulation of the Euler--Lagrange equation having those [kinematically-possible] vector fields as its family of solutions" \parencites*[292]{Curiel}.
+Ultimately, Curiel argues that Lagrangian mechanics respects the necessary and sufficient kinematic constraints that any classical system must satisfy \parencites*[307-308]{Curiel}.
+In contrast, Curiel argues that Hamiltonian mechanics does not inherently respect these kinematical constraints.
+%---interpreted as applying to those systems who possess a unique solution to the Euler--Lagrange equation---. But, Curiel is comfortable talking about Lagrangian systems that are not hyper-regular.
+
+Instead, to treat classical systems according to Hamiltonian mechanics, Curiel argues that we necessarily have to focus on a proper subset of all possible Hamiltonians.
+This subset treats conjugate momentum as a linear function of velocity: $p_i = \lambda^j_i \dot{q}_j $ \parencites*[304]{Curiel}.
+Curiel seems to take this relationship between momentum and velocity as (physically) necessary for a system to be classical.
+It entails that the Hamiltonian takes the following quadratic form---where the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q_i$:
+\begin{equation} \label{6.2}
+H =\alpha^{m n} p_m p_n + U (q_j) 
+\end{equation}
+Curiel takes this form as the ``necessary and sufficient condition [a Hamiltonian must have] for it to model an abstract classical system" \parencites*[305]{Curiel}.
+Nevertheless, Curiel notes that we can formally apply Hamiltonians that do not have the form of Equation~\ref{6.2}, e.g.
+the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$.
+Curiel then argues that all of the ``formally acceptable Hamiltonians" that do not satisfy the quadratic form (~\ref{6.2}) are neither physically significant nor physically meaningful.
+It is in this sense that Curiel believes we must artificially restrict Hamiltonian mechanics to those Hamiltonians that satisfy Equation~\ref{6.2}, in order to model physically-meaningful classical systems.
+
+We can thereby reframe Curiel's argument as follows: all physically-meaningful classical systems satisfy a certain set of kinematic constraints [expressed generally via the almost-tangent structure].
+Lagrangian mechanics encodes the necessary and sufficient conditions for a system to satisfy these kinematic constraints, both within and outside the hyper-regular domain.
+In contrast, Hamiltonian mechanics does not encode these kinematical constraints.
+Instead, it applies to non-regular systems that are not classical systems (in virtue of not satisfying these constraints).
+These systems are not physically meaningful.
+Hence, in order to make Hamiltonian mechanics physically-meaningful, we must impose an ad hoc condition (Equation~\ref{6.2}) by hand, since Hamiltonian mechanics itself does not force us to make this restriction.
+
+In response, we will defend two points.
+First, we have in fact already in Section~\ref{motion} seen a natural interpretation of both the linear relationship between momentum and velocity and the quadratic Hamiltonian form (~\ref{6.2}).
+Both follow from imposing the motion equivalence condition.
+In equation 5, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (which is connected to the gauge-redundancy of the conjugate momentum).
+Just as Curiel does, we then used this relationship to derive a quadratic form for the Hamiltonian (Equation 7).
+Imposing motion equivalence thereby allows us to interpret this quadratic form as the Hamiltonian for massive particles under potential forces.
+It is not an ad hoc condition imposed on Hamiltonian systems.
+Instead, it arises in all contexts where motion equivalence holds, i.e.
+where the trajectory of the system in physical space-time is necessary and sufficient for its evolution in state space.
+
+[To do: Is our equation 2 the same as (or intricately related to) Curiel's kinematic conditions? I.e. the case of weak equivalence.]
+
+Secondly, we believe there are some physically-meaningful non-regular Hamiltonian systems.
+These include the model of a classical photon [To do: briefly expand on this and connect to Curiel page 309-310 discussion of a similar system] We thereby view the broader scope of Hamiltonian mechanics as an advantage, rather than a disadvantage.
+Importantly, our interpretation is compatible with Curiel's claim that many---indeed, even perhaps the vast majority of---non-regular Hamiltonian systems are either physically incoherent or at least physically uninteresting.
+
+%On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems).
+In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers the Hamiltonian $H=p$, which corresponds to a system moving with constant velocity (e.g.
+velocity of 1).
+So this could very well be a model of a photon, moving at the speed of light.
+Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating the Euler--Lagrange equation.
+Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310].
+But then does Curiel consider this system to not be a classical system?
+
+More broadly, it seems that Curiel simply privileges a different definition of `classical system' than the one we adopt here.
+His definition takes satisfying his preferred kinematical constraints as fundamental.
+In contrast, we privilege a system having a unique solution, i.e.
+being deterministic and reversible.
+Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}, but he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage.
+On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}.
+In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy a particular set of physically significant kinematical constraints [(his equations 4.3)].
+Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will allegedly still satisfy the kinematical constraints of Lagrangian mechanics \parencites*[311]{Curiel}.
+Curiel therefore views these non-regular, Lagrangian but non-Hamiltonian systems as physically-meaningful classical systems.
+
+Yet, it is difficult to understand what a Lagrangian system outside the hyper-regular domain could mean.
+For outside this domain, the Euler--Lagrange equations do not have a single, unique solution.
+As an example, consider the case where the Lagrangian $L$ equals zero.
+This Lagrangian corresponds to a Hamiltonian that also equals zero, entailing that $dH = 0 $.
+By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.
+For this ``system," all paths have the same action, so all paths are equally possible.
+In contrast,  a Hamiltonian defines a unique solution (i.e.
+a unique evolution in phase space) even when outside the hyper-regular domain (i.e.
+when the Hessian is zero).\footnote{In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
+There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.}
+
+Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections.
+Additionally, we take ourselves to have responded to Curiel's main criticism of Hamiltonian mechanics, namely that it must impose \textit{ad hoc} the condition (~\ref{6.2}).
+We take these considerations as reason enough for adopting our preferred notion of a classical system.
+Of course, if one insists on defining a `classical system' as Curiel does, one will continue to have grounds to privilege Lagrangian mechanics.
+It is not our aim here to enter into a burden of proof dispute regarding which definition of `classical system' is to be strictly preferred and why.
+
+% Whereas we happily welcome at least some non-hyper-regular systems into the fold, Curiel seemingly wants to exclude these from the honorific `classical system' \parencites*[305]{Curiel}.
+Hence, Curiel views it as a disadvantage of Hamiltonian mechanics that it applies outside the hyper-regular regime, whereas we view this as a conceptual advantage.
+Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections.
+% [To do: but is the argument at the bottom of Curiel page 305 sufficiently compelling?
+Argument seems to appeal to dimensional analysis.
+He argues that Hamiltonian systems outside of this regime include ones that do not make physical sense and violate dimensional analysis.]
+
+%systems whose Hamiltonian takes the form (equation 6.2) $H =\alpha^{m n} p_m p_n + U (q_j) $.
+Here, the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q $.
+
+\begin{comment}
+
+Previous material and notes on Curiel, including additional possible responses:
+
+\begin{quotation}
+Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
+Its structures do not provide the appropriate concepts and tools to \dots provide any natural justification for the restriction to Hamiltonians of the form (6.2).
+\parencites[306]{Curiel}
+\end{quotation}
+%dots part: `formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they'
+
+% [Set this up properly: Curiel's definition of abstract classical system; Curiel's privileging of certain kinematical constraints, his equations 4.3; Hamiltonian form of equation 6.2; see below].
 Curiel restricts Hamiltonian mechanics to systems whose Hamiltonian takes the form (equation 6.2) $H =\alpha^{m n} p_m p_n + U (q_j) $.
@@
 
-\begin{quotation}
-Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
-Its structures do not provide the appropriate concepts and tools to formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they provide any natural justification for the restriction to Hamiltonians of the form (6.2).
-\parencites[306]{Curiel}
-\end{quotation}
-
 We agree with Curiel that his equation 6.2 requires a condition that is imposed by hand, namely the motion equivalence assumption [explain this].
 However, this condition has a clear physical meaning, and it corresponds to a physical condition that must hold in order to define/apply Lagrangian mechanics.
-Pace Curiel, there exist Hamiltonian systems that do not satisfy his equation 6.2 (equivalently, classical systems where the momentum is not a linear function of the velocity), and these will even have a Lagrangian, although they will not be an abstract classical system in Curiel's sense.
+\textit{Pace} Curiel, there exist Hamiltonian systems that do not satisfy his equation 6.2 (equivalently, classical systems where the momentum is not a linear function of the velocity), and these will even have a Lagrangian, although they will not be an abstract classical system in Curiel's sense.
 Hence, there seems to be a burden on Curiel to distinguish---in a principled way---the classical Lagrangian systems from those that are not classical.
-% Hence, perhaps we agree with Curiel that ``it is not necessary that [these conditions] hold in order for the theory to be meaningfully applied to model a [classical] system" [305].
-Where is Curiel takes this as an argument for the primacy of Lagrangian mechanics, we take it as an argument for the primacy of Hamiltonian mechanics, since Curiel's conditions are equivalent to imposing motion equivalence.
+% Hence, we agree with Curiel that ``it is not necessary that [these conditions, Eqn 6.2] hold in order for the theory to be meaningfully applied to model a [classical] system" [305].
+Whereas Curiel takes this as an argument for the primacy of Lagrangian mechanics, we take it as an argument for the primacy of Hamiltonian mechanics, since Curiel's conditions are equivalent to imposing motion equivalence.
 
@@
 
+Figure out how to relate the following to the preceding: In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
+There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.
+Indeed, outside the hyper-regular domain, the Euler--Lagrange equations do not have a single, unique solution.
+Hence, it is difficult to say what a Lagrangian system even is outside the hyper-regular domain.
+As an example, consider the case where the Lagrangian $L$ equals zero.
+Then for this ``system," all paths have the same action, so all paths are equally possible.\footnote{This Lagrangian corresponds to a Hamiltonian that also equals zero, entailing that $dH = 0 $.
+By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.} In contrast,  a Hamiltonian defines a unique solution (unique evolution in phase space) even when outside the hyper-regular domain (i.e.
+when the Hessian is zero).
+
 Summary: Hamiltonian mechanics provides solutions outside hyper-regular domain (but not these kinematical constraints) whereas Lagrangian mechanics preserves these kinematical constraints outside hyper-regular domain.
@@
 
+\end{comment}
+
+\section{Conclusion}
+
+To do!
+
+\begin{comment}
+
 \section{Material looking for a possible home}
@@
 Teh and Tsementzis \parencites*[]{Teh} provide another argument for the equivalence of Lagrangian and Hamiltonian mechanics within the hyper-regular domain.
-They develop this using Tulczyjew's reformulation of Hamiltonian mechanics, which shows that ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh}.
+They develop this using Tulczyjew's reformulation of Hamiltonian mechanics, which shows that ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh} [I slipped in part of this claim as a footnote into section 2.1: shows another reason to view Hamiltonian mechanics as fundamental in a non-metaphysical sense].
+Rest of the discussion seems too complicated and technically involved to set up properly.
+Response-claims below might not even be completely accurate.
 Again, there is a worry that this equivalence result might undercut our privileging of Hamiltonian mechanics.
@@
 [To do: Not sure where to put the following material, if anywhere; perhaps after a discussion of why we restrict to hyper-regular domain?:]
-In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
-There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.
-Indeed, outside the hyper-regular domain, the Euler--Lagrange equations do not have a single, unique solution.
-Hence, it is difficult to say what a Lagrangian system even is outside the hyper-regular domain.
-As an example, consider the case where the Lagrangian $L$ equals zero.
-Then for this ``system," all paths have the same action, so all paths are equally possible.\footnote{This Lagrangian corresponds to a Hamiltonian that also equals zero, entailing that $dH = 0 $.
-By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.} In contrast,  a Hamiltonian defines a unique solution (unique evolution in phase space) even when outside the hyper-regular domain (i.e.
-when the Hessian is zero).
-
-\begin{comment}
 

### 2021-02-02 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex

@@
 For systems that satisfy the motion equivalence condition, then, the following relationships hold at any given time:
-\begin{equation}
+\begin{equation}\label{weak_equivalence}
 \begin{aligned}
@@
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).
-Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.
+Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.\footnote{Weak equivalence corresponds to a general treatment of Lagrangian mechanics that does not presuppose Riemannian geometry.}
 
@@
 distributions on a subset of the coordinates), which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
-We set:
+We set:\footnote{In Section~\ref{dependence}, we will see that $g^{ij}$ corresponds to the Riemannian metric, since $\partial v^i/\partial p_j$ it is a function only of position.}
 \begin{equation}
@@
 At first glance, this expression misleadingly suggests that the density $\hat{\rho}$ is coordinate-independent: it is a scalar quantity with the correct dimensions of mass over volume.
-[XXX to do: ask GC if the following explanation is accurate] Nevertheless, $\hat{\rho}$ remains coordinate-dependent because specifying a numerical value for $\hat{\rho}$ requires choosing a particular set of coordinates.
-Fortunately, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
+However, on physical grounds, we require that the numerical value of $\hat{\rho} $ equals the value of $\rho$.
+This is only possible if the determinant of $g_{ab}$ equals one.
+Hence, we must define $\hat{\rho} $ in a special coordinate system, namely one where---at least locally---$|g_{ab}| = 1$.
+For this reason, $\hat{\rho} $ remains coordinate-dependent.
+%Fortunately, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
 Since the square root of the determinant $|g_{ab}|$ is the Jacobian of the transformation to those coordinates, $|g_{ab}|$ will be constant only when this Jacobian equals one.
-For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$; in polar coordinates $|g_{ab}| = r^2$ and hence the volume form $dV = r dr \wedge d\theta$.
+
+For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$.
+Whereas in polar coordinates, $|g_{ab}| = r^2$ and hence the volume form $dV = r dr \wedge d\theta$.
 This means that $\hat{\rho}$ at best expresses the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise.\footnote{This is similar to what happens with the concept of proper time: it is defined in the co-moving frame, so it is coordinate dependent.
@@
 For instance, in her argument for the fundamentality of Hamiltonian mechanics, North claims that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
-[To do: see if Wallace 2003 also makes this claim about the fundamentality of conjugate momentum, or the similarly flawed claim that conjugate momentum and position are on equal footing within Hamiltonian mechanics].
+[To do: incorporate Wallace 2003 (p.
+164) claim that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
+See Curiel 2014, ft 16].
 Yet, as we show below, the conjugate momentum $p$ is actually a gauge-redundant quantity.
@@
 He interprets this as capturing the fundamental kinematical constraints relating position and velocity in classical mechanics.} Central to Curiel's argument is the claim that Lagrangian mechanics naturally encodes the kinematical structure of classical systems whereas Hamiltonian mechanics requires imposing an \textit{ad hoc} condition.
-We will now argue that the condition Curiel claims is `ad hoc' actually has a natural physical interpretation.
-Additionally, we note a disagreement about which formal classical systems are physically significant.
+We will now argue that the condition Curiel claims is ad hoc actually has a natural physical interpretation.
+Additionally, we consider a further disagreement about the proper definition of classical systems, which bears on which formal models are interpreted as physically significant.
+% Additionally, we provide reasons to prefer our definition of classical systems over Curiel's, by considering classical systems outside the hyper-regular domain.
 
@@
 In contrast, Curiel argues that Hamiltonian mechanics does not inherently respect these kinematical constraints.
+Essentially, the kinematical constraints are given by velocity being the first temporal derivative of position: $v^i = d_t x^i = v^i(q^j, p_k)$ (equation~\ref{weak_equivalence}).\footnote{Curiel formalizes this kinematical constraint using the geometrical notion of an almost-tangent structure \parencites*[290]{Curiel}.} Curiel contends that this constraint is `kinematical' in the sense that it must hold independently of the interactions between a system and its environment.
+He interprets it as a substantive physical condition, rather than an analytic truth \parencites*[282]{Curiel}.
 %---interpreted as applying to those systems who possess a unique solution to the Euler--Lagrange equation---. But, Curiel is comfortable talking about Lagrangian systems that are not hyper-regular.
@@
 the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$.
-Curiel then argues that all of the ``formally acceptable Hamiltonians" that do not satisfy the quadratic form (~\ref{6.2}) are neither physically significant nor physically meaningful.
+Curiel then argues that all of the ``formally acceptable Hamiltonians" that do not satisfy the quadratic form (\ref{6.2}) are neither physically significant nor physically meaningful.
 It is in this sense that Curiel believes we must artificially restrict Hamiltonian mechanics to those Hamiltonians that satisfy Equation~\ref{6.2}, in order to model physically-meaningful classical systems.
 
-We can thereby reframe Curiel's argument as follows: all physically-meaningful classical systems satisfy a certain set of kinematic constraints [expressed generally via the almost-tangent structure].
+We can thereby reframe Curiel's argument as follows: all physically-meaningful classical systems satisfy a certain set of kinematic constraints, e.g.
+$v^i = d_t x^i$.
 Lagrangian mechanics encodes the necessary and sufficient conditions for a system to satisfy these kinematic constraints, both within and outside the hyper-regular domain.
@@
 Instead, it applies to non-regular systems that are not classical systems (in virtue of not satisfying these constraints).
-These systems are not physically meaningful.
-Hence, in order to make Hamiltonian mechanics physically-meaningful, we must impose an ad hoc condition (Equation~\ref{6.2}) by hand, since Hamiltonian mechanics itself does not force us to make this restriction.
+According to Curiel, these systems are not physically meaningful.
+Hence, in order to make Hamiltonian mechanics physically-meaningful, we must impose an ad hoc condition by hand (Equation~\ref{6.2}), since Hamiltonian mechanics itself does not force us to make this restriction.
 
 In response, we will defend two points.
-First, we have in fact already in Section~\ref{motion} seen a natural interpretation of both the linear relationship between momentum and velocity and the quadratic Hamiltonian form (~\ref{6.2}).
+First, we have in fact already in Section~\ref{motion} seen a natural interpretation of both the linear relationship between momentum and velocity and the quadratic Hamiltonian form of equation~\ref{6.2}.
 Both follow from imposing the motion equivalence condition.
@@
 
-[To do: Is our equation 2 the same as (or intricately related to) Curiel's kinematic conditions? I.e. the case of weak equivalence.]
+% Note that our equation 2 (\ref{weak_equivalence}) is a special case (or less mathematically sophisticated formulation) of Curiel's kinematic conditions, i.e.
+his equation 4.3 in terms of the almost-tangent structure.
+Our equation 2b matches Curiel's equations 3.1a and 3.2b, in the case of free and non-free Newtonian particles, respectively.
 
 Secondly, we believe there are some physically-meaningful non-regular Hamiltonian systems.
-These include the model of a classical photon [To do: briefly expand on this and connect to Curiel page 309-310 discussion of a similar system] We thereby view the broader scope of Hamiltonian mechanics as an advantage, rather than a disadvantage.
-Importantly, our interpretation is compatible with Curiel's claim that many---indeed, even perhaps the vast majority of---non-regular Hamiltonian systems are either physically incoherent or at least physically uninteresting.
+These include models of a classical photon.
+In general, Hamiltonian systems outside the hyper-regular domain correspond to cases where we cannot express the density of states in terms of position and velocity.
+These are cases where the weak equivalence condition (equation~\ref{weak_equivalence}) fails.
+Nevertheless, we can still define a distribution over the phase space points and consider its evolution.
+We thereby view the broader scope of Hamiltonian mechanics as an advantage, rather than a disadvantage.
+This interpretation is compatible with the claim that many---perhaps even most---non-regular Hamiltonian systems are physically uninteresting.
+%[To do: briefly expand on this and connect to Curiel page 309-310 discussion of a similar system]
 
@@
 But then does Curiel consider this system to not be a classical system?
+Additionally, insofar as this corresponds to the trivial Lagrangian, does this pose a problem for his own endorsement of Lagrangian mechanics outside the hyper-regular domain?
 
 More broadly, it seems that Curiel simply privileges a different definition of `classical system' than the one we adopt here.
-His definition takes satisfying his preferred kinematical constraints as fundamental.
+His definition takes satisfying certain kinematical constraints as fundamental.
 In contrast, we privilege a system having a unique solution, i.e.
@@
 On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}.
-In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy a particular set of physically significant kinematical constraints [(his equations 4.3)].
-Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will allegedly still satisfy the kinematical constraints of Lagrangian mechanics \parencites*[311]{Curiel}.
+In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy the proper kinematical constraints (e.g.
+$v^i = \dot(x)^i$).
+Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will allegedly still satisfy these kinematical constraints \parencites*[311]{Curiel}.
 Curiel therefore views these non-regular, Lagrangian but non-Hamiltonian systems as physically-meaningful classical systems.
@@
 By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.
-For this ``system," all paths have the same action, so all paths are equally possible.
+For this ``system," all paths have the same action, so all paths are possible.
+This gives an initial probability distribution over the states, but it does not specify how this initial distribution will evolve into a future distribution of states.
+In other words, the Lagrangian alone does not supply a probability distribution on the possible evolutions.
 In contrast,  a Hamiltonian defines a unique solution (i.e.
@@
 
-Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections.
-Additionally, we take ourselves to have responded to Curiel's main criticism of Hamiltonian mechanics, namely that it must impose \textit{ad hoc} the condition (~\ref{6.2}).
+By providing a clearer conception of classical systems outside the hyper-regular domain, we believe that our definition of a `classical system' has advantages over Curiel's.
+It provides a uniform treatment of phase space distributions, accommodates models of a classical photon, and maintains uniqueness of solutions.
+For this reason, we believe that the burden is on Curiel to clarify the nature of non-hyper-regular Lagrangian systems.
+Additionally, we take ourselves to have responded to Curiel's main criticism of Hamiltonian mechanics, namely that it must impose \textit{ad hoc} the condition (\ref{6.2}).
 We take these considerations as reason enough for adopting our preferred notion of a classical system.
 Of course, if one insists on defining a `classical system' as Curiel does, one will continue to have grounds to privilege Lagrangian mechanics.
-It is not our aim here to enter into a burden of proof dispute regarding which definition of `classical system' is to be strictly preferred and why.
+Although we have given reasons to prefer our definition of a classical system, these considerations may not be decisive.\footnote{As described elsewhere, our definition of a classical system has the advantage of an elegant relationship with quantum mechanical systems.
+By changing the requirement of infinitesimal reducibility (Section~\ref{infinitesimal}) to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.
+Hence, our definition elegantly distinguishes between quantum and classical systems.
+Both kinds of systems are Hamiltonian thanks to deterministic and reversible evolution.} Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections.
+
+%  It is not our aim here to enter into a burden of proof dispute regarding which definition of `classical system' is to be strictly preferred and why.
 

### 2021-02-02 Articles/2019-ShannonEntropy/Source/ShannonEntropyIsVariability.tex
Preparing new arxiv submission

@@
-%\documentclass[prb,preprint]{revtex4-1}
+\documentclass[nofootinbib,twocolumn]{revtex4-1} 
+% The line above defines the type of LaTeX document.
+% Note that AJP uses the same style as Phys. Rev. B (prb).
 
-\documentclass{article}
-\textheight 8.9 in \textwidth 6.5 in \oddsidemargin -10pt \topmargin -30pt
+% The % character begins a comment, which continues to the end of the line.
 
-\usepackage{amsmath}
-%\usepackage[usenames,dvipsnames]{color}
-
-\title{Universal Characterization of Physical Variability by \\Shannon's $-\sum_ip_i\,\textrm{ln}\,p_i$ Formula}
-
-\author{
-	Gabriele Carcassi,
-	Julian Barbour and
-	Christine A. Aidala 
-}
+\usepackage{amsmath}  % needed for \tfrac, \bmatrix, etc.
+\usepackage{amsfonts} % needed for bold Greek, Fraktur, and blackboard bold
+\usepackage{graphicx} % needed for figures
+\usepackage{hyperref}
 
 \begin{document}
+
+\title{Variability as a better characterization of Shannon entropy}
+
+\author{Gabriele Carcassi}
+\email{carcassi@umich.edu}
+\author{Christine A. Aidala}
+\email{caidala@umich.edu}
+\affiliation{Department of Physics, University of Michigan, Ann Arbor, MI 48109-1040 }
+% Please provide a full mailing address here.
+
+\author{Julian Barbour}
+\email{julian.barbour@physics.ox.ac.uk}
+\altaffiliation[permanent address: ]{College Farm, South Newington, Banbury, Oxon, OX15 4JG UK}
+\affiliation{Visiting Professor in Physics at the University of Oxford (2009-2018)}
+
+\date{\today}
+
+\begin{abstract}
+	The Shannon entropy, one of the cornerstones of information theory, is widely used in physics, particularly in statistical mechanics.
+Yet its characterization and connection to physics remain vague, leaving ample room for misconceptions and misunderstanding.
+We will show that the Shannon entropy can be fully understood as measuring the variability of the elements within a given distribution: it characterizes how much variation can be found within a collection of objects.
+We will see that it is the only indicator that is continuous and linear, that it quantifies the number of yes/no questions (i.e.~bits) that are needed to identify an element within the distribution, and we will see how applying this concept to statistical mechanics in different ways leads to the Boltzmann, Gibbs and von Neumann entropies.
+\end{abstract}
 
@@
 
-\begin{abstract}
-TBD
-\end{abstract}
-
 \section{Introduction\label{int}}
 
-The use of Shannon entropy $-\sum p_i \log p_i$ is well established in very different fields of knowledge such as computer science[cite], physics[cite] and ecology[cite].
-Within computer science it is a fundamental element in information theory since Shannon developed his mathematical theory of communication.
-It has been used extensively within statistical mechanics to link the microscopic description of a system to its macroscopic thermodynamic one.
-It has been used in ecology to characterize biodiversity and calculate species distributions under given constraints.
-Entropy maximization has been shown to be a powerful tool.
+The goal of this work is to give a comprehensive introduction to the concept of Shannon entropy, the expression $-\sum p_i \log p_i$, which is well established in such disparate fields as computer science, communication theory, physics \cite{Jaynes1957-1, Grandy} ecology and economics.
+The concept of entropy in physics, despite being treated extensively in the physics education literature (e.g.
+\cite{baierlein1994entropy, carson2002undergraduate, Santillan_2008, LeffPartI, LeffPartII, LeffPartIII, LeffPartIV, LeffPartV, Benguigui_2013, geller2014entropy, Tiwary_2020}), is still surrounded by a great deal of confusion \cite{Swendsen, Styer}.
+Partly because, at a fundamental level, there is still disagreement on what is the ``correct'' notion of entropy, particularly out of equilibrium \cite{jarzynski2017stochastic, goldstein2019gibbs, maes2003time}.
+Partly because the same word ``entropy'' is used to refer to concepts (e.g.
+topological entropy, thermodynamic entropy, graph entropy, R\'{e}nyi entropy, information entropy) that have different definitions and various degrees of overlap.
+We will focus on the notion of Shannon entropy, by which we mean the use of $- \sum_i p_i \log p_i$ in the broadest characterization, which has received less attention in physics education.
+We will see that the Shannon entropy can be given a crisp and precise characterization and will explore its relation to the Gibbs and Boltzmann entropies in statistical mechanics.\footnote{We will not discuss its relation to thermodynamic entropy.
+Thermodynamic entropy is defined on the notions of heat and work while the Shannon entropy is defined on the notion of a distribution.
+The contexts are different and relating them goes beyond the scope of this work.
+} The same material is also presented in a series of videos \cite{videos}.
 
-While the techniques are well developed, it is puzzling that this concept can be applied in such different contexts, given that it assumes different features in each one.
-For example, in information theory the Shannon entropy for a book will change whether we choose to encode words or letters, which means the quantity does not have a unique value: it depends on an arbitrary choice.
-This is not problematic for an information theoretic quantity, but it is for a physical quantity.
-In statistical mechanics the $p_i$ represent probabilities, which means the Shannon entropy is a property of an ensemble of physical systems.
-This leads some to claim that the Shannon entropy is epistemic in nature and represents ignorance.
-[Add references] In ecology, however, the $p_i$ represent the fraction of individuals belonging to the $i$th species.
-There is no probabilistic ensemble to which one can associate a degree of ignorance.
-Finally, entropy in physics is associated to the notion of states and irreversible time evolution, which are not well defined concepts for the other fields.
-The same expression, then, seems to have different traits and it is not clear why a single concept would apply in all these cases.
+To give a crisp and intuitive meaning to the Shannon entropy we need to address two main problems.
+The first problem is a consequence of its success.
+As it is used in many fields, it is generally introduced with concepts that are specific to that field but may not be appropriate for others.
+As we will see, the meaning of the $p_i$ can vary significantly.
+The first goal, then, is to provide a characterization of the Shannon entropy that satisfies the following requirements:
+\begin{itemize}
+	\item it is defined in a general setting that is crisp and intuitive
+	\item it is precise enough so that the formula itself can be derived
+	\item it recovers the specific meaning when applied to each field.
+\end{itemize}
+In this paper we show that the idea of variability of the elements within a distribution, how diverse the objects are from each other within the collection, can serve as such a general concept.
+We will therefore use the term \emph{``Shannon variability''} for this more general idea that applies to all cases, leaving ``entropy'' to the more specific physical concept.
+We will show that the expression $- \sum_i p_i \log p_i$ is the only linear indicator of variability and it measures the number of questions one must ask to identify an element of the distribution, linking its use to information theory.
+The variability of messages within a communication channel will correspond to its information entropy as used in communication theory; the variability of microstates within a macrostate will correspond to the Gibbs entropy used in statistical mechanics; the variability of the state of a single particle within a microstate will correspond to the Boltzmann entropy.
+This avoids the use of incomplete or imprecise characterizations (e.g.
+uncertainty, knowledge, lack of knowledge, disorder) that, while useful in some limited circumstances, promote confusion in others.
 
-To make matters worse, it is fairly typical to describe entropy with notions such as uncertainty[cite], knowledge[cite], lack of knowledge[cite], disorder[cite].
-While some of these characterizations are useful in some contexts to give an intuitive picture, they do not always seem to work.
-A common objection to the characterization as disorder, for example, is that a uniform distribution corresponds to a homogeneous system, which one would expect to be maximally ordered; yet the distribution maximizes the Shannon entropy, which would indicate maximal disorder.
-Moreover, these notions remain vague in the sense that the formula is not formally derived from those concepts, and therefore it is not clear why that particular indicator would be an appropriate choice.
-We are left to wonder: is there really some single underlying concept, that, when applied appropriately in the different fields, reduces to the correct notion of entropy?
-Or is it just a big coincidence?
+The second problem is that the Shannon formula presents some peculiarity when applied over continuous variables that is often the source of confusion.
+Namely, it is generally linked to a choice of unit, of the resolution used to describe the problem.
+However, the physical state spaces, in both classical and quantum mechanics, provide additional structure that make the Shannon variability coordinate-independent.
+This is seldom pointed out, and it is critical to form a proper connection to the Boltzmann, Gibbs and von Neumann entropies.
 
-We propose that there is indeed a characterization of Shannon's formula that is both more general and more appropriate; that is independent of whether it is applied to sources of messages, ensembles of microstates or animal populations.
-We will argue that the formula is a measure of the \emph{variability of the values within a distribution of a given set of elements}.
-That is, given a distribution of some elements over some quantity, the Shannon entropy represents how varied, how different from each other, the values are.
-In fact, it is the only continuous and linear index that measures the degree of variability within a distribution, and those requirements lead naturally to the same three conditions that Shannon set for his formula.\cite{Shannon} In other words, the concept of variability within a distribution leads in a formal way to the correct expression.
-In all cases where the formula is applied, in fact, we are using distributions as objects of our investigation and the variability within that distribution will be of interest.
-Yet, the physical meaning of those distributions can be widely different: they may represent the fractional distribution of animals according to their species; they may represent probability distributions of different outcomes; they may represent knowledge of certain facts as in Bayesian statistics.
-Therefore the variability itself, in the context of these different cases, will necessarily acquire a different meaning.
-In particular, we will see how this perspective naturally unfolds in the contexts of information theory and statistical mechanics.
-
-As we believe this is the right characterization, we will therefore use the expression Shannon ``variability'', rather than ``entropy'', ``information'' or ``diversity index'', as it tells us exactly what the quantity describes.
-It is well known that when Shannon recognized the importance of his formula in communication theory he sought von Neumann's advice on the name he should give to the expression he had found.
-He was told to call it \emph{entropy}, first, because Boltzmann had used an identical formula to define entropy in statistical mechanics and, second, nobody understood entropy and that would provide protection from criticism [exact words to be found].
-We believe von Neumann's advice was not totally helpful and has given rise to confusion.
-We will reserve the word ``entropy'' for the one dealing with physical states and irreversible processes.
-
-This characterization also explains why the Shannon variability has such a broad range of applicability.
-As nature presents itself in a multitude of forms and combinations, a fundamental task in science is to recognize which objects belong together and study how their different attributes are expressed in different ways or degrees.
-They may be galaxies, animals or outcomes of the same process; we may study their shape, behavior or numerical quantities.
-But the idea of grouping them and trying to understand the source of their variation is a common endeavor in science.\footnote{In fact, determining which collection of objects is significant and what attributes are relevant is a significant part of the work.} Measuring and explaining this variability is an essential task.
-It is this very variety that is the inspiration for and the subject of scientific investigation.
-A universe without variety, where all objects were the same and remained the same, is barely conceivable and certainly it would leave no space for scientific investigation.
-In the words of Leibniz, the principle underlying the universe is that it should realize ``as much variety as possible, but with the greatest order possible''.\footnote{MOVE TO BIBLIO Paragraph 58 of Leibniz's \emph{Mondadology}}
-
-As a final note, we want to stress that the goal of this work is not to present some new technical result.
-The goal is instead to synthesize all details into a coherent picture that gives them context and a precise relationship to each other.
-This can help, especially the new practitioners, to avoid common misconceptions.
-Moreover, while some details presented here may be well known in a particular community, we find that they may be new to another.
-For instance, we will show an example of Huffmann coding to illustrate the meaning of the numerical value associated with the Shannon variability.
-This is standard practice in computer science courses, yet it is not in statistical mechanics.
-Therefore restating all the crucial aspects in one place is beneficial to make the connection between them readily available.
-
-%https://www.researchgate.net/publication/259011656_A_tribute_to_Claude_Shannon_1916-2001_and_a_plea_for_more_rigorous_use_of_species_richness_species_diversity_and_the_'Shannon-Wiener'_Index
-
-%https://web.stanford.edu/~hastie/Papers/maxent_explained.pdf
-
-%https://www.mdpi.com/1099-4300/21/8/794/htm
-
-%https://www.sciencedirect.com/science/article/abs/pii/S0169534714001037
-
-%Leibniz
-%We recall, though we have been unable to relocate, a sentence in one of Leibniz's many writings to the effect that reality \emph{is} variety.
-Whether he made the statement or not, it is consistent with his view that any contingent thing is defined by a collection of attributes, which necessarily must be distinguishable at some epistemic level.
-Paragraph 58 of Leibniz's \emph{Mondadology} asserts that the principle underlying the universe is that it should realize ``as much variety as possible, but with the greatest order possible''.
-[REF]
-
-%We conclude this section with a word on terminology.
-In mathematics, Cantor defined a set as ``a collection into a whole $M$ of definite and separate objects''.
-He called the objects \emph{elements}.
-We will also use the word `elements' to denote the things we consider but subject to the proviso that they have a distinguishing feature (or features) that allows us to say that \emph{they are of the same kind}.
-Thus, we consider sets of a restricted type.
-The point is this.
-To constitute a set in accordance with Cantor's definition, it is only necessary for its elements to be \emph{definite and separate} objects.
-This leaves total freedom as to how such objects in the universe are to be parcelled up into specific sets.
-For example, a set might consist of Mars and the word `go'.
-We think such latitude misses one of the most significant aspects of the universe, namely, that the things within it can be distinguished by attributes that are much more specific than the two required by Cantor.
-Our field of investigation goes beyond mere number, which washes out all attributes apart from the two, of being definite and separate, that allow us to count.
-They were the only attributes that Cantor needed; we need more.
+While some details presented here may be well known in a particular community, we find that they may be new to another.
+Therefore we assume little knowledge of the subject from the reader and include all details that we feel are important to properly understand the subject without confusion, including two standard derivations for the formula and examples that are common in computer science but rarely used in physics.
+We will briefly touch on a few of the common misconceptions and show how they are resolved.
 
@@
 Once we arbitrarily decided we will study the galaxies in the Laniakea Supercluster, it is a matter of fact that the Milky Way will be included.
-This also means that, whatever properties those elements will have, they will represent a matter of fact about that set.
-In turn, this will make some choices arguably more appropriate or interesting than others.
+This also means that, whatever properties those elements have, they will represent a matter of fact about that set.
+In turn, this will make some choices arguably more appropriate or insightful than others.
 For example, in taxonomy the set of all animals that have feathers may be more appropriate than the set of all animals that can fly: the first includes only birds while the second includes most birds, a few mammals and a lot of insects.
@@
 
-As we noted before, the level at which we describe the elements is not absolute and can change.
-For example, we may choose to group the animals in the Gal\'{a}pagos islands first by class (e.g.
-mammals, birds, reptiles) and then later refine the mammals by species.
+As noted before, the level at which we describe the elements is not absolute and can change.
+For example, we may choose to group the animals in the Gal\'{a}pagos islands first by class (e.g.~mammals, birds, reptiles) and then later refine the mammals by species.
 In this case, we would like $H$ to combine linearly with respect to the weights.
@@
 \end{enumerate}
-We note that these are exactly the requirements Shannon put forth for his expression\cite{Shannon}, and that these requirements lead to only one possible option: $H(p_i) = - \sum p_i \log(p_i)$.
-
-The Shannon variability is a clear and important attribute of a given distribution: it quantifies the variability of the elements within the distribution, the variety of values one finds.
+These are the same requirements Shannon put forth for his expression\cite{Shannon}, from which he showed that the only possible choice is $H(p_i) = - \sum p_i \log(p_i)$.
+See also \cite{khinchin2013mathematical} for a similar derivation.
+This expression quantifies the variability of the elements within the distribution, the variety of values one finds.
 The precise meaning of this variability is context dependent, as the choices of the elements and binning are not fixed and the meaning of the weights depends on what the distribution is describing.
 But this is true for any mathematical object: a real number may represent mass, color in the frequency spectrum, the total money supply, the half-life for an isotope, a probability and so on.
-And like a real number represents a quantity, transcending what is physically quantified, the Shannon variability represents the variability of the elements in a distribution, transcending what the distribution physically signifies.
 
@@
 \begin{enumerate}
-\item There is 50\% chance you won one million dollars in the lottery and 50\% chance you didn't.
+	\item There is 50\% chance you won one million dollars in the lottery and 50\% chance you won nothing.
 	\item There is 50\% chance you won one million dollars in the lottery and 50\% chance you won half a million dollars.
 \end{enumerate}
-The distribution in both cases is the same, two bins 50\% chance each, and so is the Shannon variability. Yet, you know more in the second case: you know you have won at least half a million dollars.
+The distribution in both cases is the same, two bins 50\% chance each, and so is the Shannon variability. Yet, you know more in the second case: you know you won at least half a million dollars.
 
@@
 If we look at the galaxies in our universe, what variability do they exhibit in terms of their types?
-If we look at the animals in the Gal\'{a}pagos island, what variability is expressed in terms of their species?
+If we look at the animals in the Gal\'{a}pagos islands, what variability is expressed in terms of their species?
 If we look at the molecules in a given macrostate, what variability do they express in terms of their microstates?
 If we look at the words in the complete works of Shakespeare, what variability do we find in his vocabulary?
-All these questions are interesting, all these questions are insightful and the answers to all these questions are expressed through just one thing: the Shannon variability.
 
-If the Shannon expression is a measure of variability, how is it connected to information? What is it measuring and in what units?
+If the Shannon expression is a measure of variability, why is it connected to information? How is variability quantified and in what units?
 
@@
 
-We have seen that the Shannon variability measures the variability within a distribution but, in the end, it is a numeric quantity and we cannot expect to understand the concept if we do not understand what the number represents and in what units.
-
-We can gain insight with the following example.
+To understand what the numerical value represents, consider this example.
 Suppose we fix a distribution, say the animals in the Galap\'{a}gos islands binned by their respective species.
@@
 Not all questions will be able to extract the same amount of information.
-Some questions, like ``is it an animal?" would be redundant.
-Others, like ``is it an American Flamingo (Phoenicopterus ruber)?", would give us a lot of information in the positive case but little in the negative case.
-However, it is clear that there is always a minimum number of questions that must be asked to get to the answer.
-Given that there are more than two species, a single question cannot be enough.
-It should however be intuitively clear that to a greater variability within the distribution will correspond a greater number of questions you must ask.
+Some questions, like, ``Is it an animal?" would be redundant.
+Others, like, ``Is it an American Flamingo (Phoenicopterus ruber)?", would give us a lot of information in the positive case but little in the negative case.
+However, there has to be a minimum number of questions that must be asked to get to the answer.
+A single question, for example, cannot be enough given that there are more than two species.
+It should be intuitively clear that to a greater variability within the distribution will correspond a greater number of questions you must ask.
 That is exactly what the Shannon variability quantifies: the minimum average number of questions one has to ask to identify a value in the distribution.
@@
 If we have binary questions, the logarithms will be in base two and the unit for Shannon variability will be bits.
-It will indicate the average number of yes/no questions we need to identify an element within the dsitribution.
+It will indicate the average number of yes/no questions we need to identify an element within the distribution.
 In general, you can pick any base $b$: in base three we have ternary questions and trits, for ten we have questions with ten possible answers and digits.
 We can also pick a non-integer base, like the natural base for logarithms, and we will have nats.
-This is why the Shannon variability is fundamental in information theory, because it quantifies how much information is needed to transfer a value from a distribution.
+This is why the Shannon variability is fundamental in information theory, because it quantifies \emph{how much information is needed to transfer a value picked from a known distribution}.
 
 Defining a set of questions means choosing an encoding as we are choosing how the information gets codified into our series of bits.
-To understand how this works, we can briefly review the Huffman coding,[CITE] which is the optimal algorithm for symbol-by-symbol coding with a known probability distribution.
+To understand how this works, we can briefly review the Huffman coding,\cite{Huffman} which is the optimal algorithm for symbol-by-symbol coding with a known probability distribution.
 The idea is that we want all possible answers to each question to be balanced, to provide the same amount of information.
@@
 For example, suppose the population of pets in a country is as follows:
-\begin{itemize}
-\item dogs 27\%
-\item cats 48\%
-\item fish 10\%
-\item bird 8\%
-\item small mammal 4\%
-\item reptile 3\%
-\end{itemize}
+\begin{table}[h]
+	\begin{tabular}{lr}
+		dogs & 27\% \\
+		cats & 48\% \\
+		fish & 10\% \\
+		birds & 8\% \\
+		small mammals & 4\% \\
+		reptiles & 3\%
+	\end{tabular}
+\end{table}
+
+\noindent
 For the first question, we group cats on one side and everything else on the other, to form a 48/52 split.
-So we can ask ``is it a cat?".
+So we can ask, ``Is it a cat?".
 If the answer is yes, we finished.
@@
 We can group dogs on one side and everything else on the other to form a 27/25 split.
-So we can ask `` is it a dog?".
+So we can ask, ``Is it a dog?".
 If the answer is yes, we are done.
 If not, we continue.
-We can group fish with reptile and bird with small mammal to form a 13/14 split.
-So we can ask ``is it a fish or a reptile?" If the answer is yes, the followup question would be ``is it a fish?" If the answer is no, the followup question would be ``is it a bird?" With this scheme, the encoding, where 1 represents 'yes' and 0 represents 'no' to the each of the questions asked, is as follows:
+We can group fish with reptiles and birds with small mammals to form a 13/12 split.
+So we can ask, ``Is it a fish or a reptile?".
+If the answer is yes, the followup question would be ``Is it a fish?".
+If the answer is no, the followup question would be ``Is it a bird?".
+With this scheme, the encoding, where 1 represents 'yes' and 0 represents 'no' to the each of the questions asked, is as follows:
 \begin{itemize}
@@
 	\item fish 10\% - 4 questions (i.e.~4 bits) - answers: [no, no, yes, yes] (i.e.~encoding 0011)
-\item bird 8\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, yes] (i.e.~encoding 0001)
-\item small mammal 4\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, no] (i.e.~encoding 0000)
-\item reptile 3\% - 4 questions (i.e.~4 bits) - answers: [no, no, yes, no] (i.e.~encoding 0010)
+	\item birds 8\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, yes] (i.e.~encoding 0001)
+	\item small mammals 4\% - 4 questions (i.e.~4 bits) - answers: [no, no, no, no] (i.e.~encoding 0000)
+	\item reptiles 3\% - 4 questions (i.e.~4 bits) - answers: [no, no, yes, no] (i.e.~encoding 0010)
 \end{itemize}
@@
 These are taken from $I$ different cases: dogs, cats, fish and so on.
-Suppose $N_i$ are the of elements of each type, which means $\sum_{i \in I} N_i = N$.
+Suppose $N_i$ are the number of elements of each type, which means $\sum_{i \in I} N_i = N$.
 Then, given a particular instance of $N_i$, we have $W(N_i) = \frac{N!}{\prod_{i \in I} N_i!}$ possible ways to realize that sequence, which corresponds to all possible permutations.
 If all permutations are equally likely, then $\log W (N_i)$ represents the number of bits needed to identify one of the sequences.
-Assuming we use base $b$ for the logarithm, we take $N$ to be large and, using Stirling's approximation $\ln N!
-\approx N \ln N - N$, we find:
-\begin{equation}
-\begin{aligned}
-\log W(N_i) &= \log \frac{N!}{\prod_{i \in I} N_i!} \\
-&= \log N! - \sum_{i \in I} \log N_i! \\
-&\approx \frac{1}{\ln b} \left(N \ln N - N - \sum_{i \in I} N_i \ln N_i + \sum_{i \in I} N_i \right) \\
-&= \frac{1}{\ln b} \left(\sum_{i \in I} N_i \ln N - \sum_{i \in I} N_i \ln N_i \right) \\
-&= -\sum_{i \in I} N_i \left(\log N_i - \log N \right) \\
-&= - N \sum_{i \in I} \frac{N_i}{N} \log \frac{N_i}{N} \\
-&= - N \sum_{i \in I} p_i \log p_i \\
-&= N H(p_i)
-\end{aligned}
-\end{equation}
 
+When $N$ is large, we can use Stirling's approximation $\ln N!
+\approx N \ln N - N$, and find $\log W(N_i) \approx - N \sum_{i \in I} \frac{N_i}{N} \log \frac{N_i}{N} = N H(p_i)$.
+The logarithm of the permutations is $N$ times the Shannon variability.
 The result should not be surprising: it simply tells us that encoding a sequence of $N$ elements is the same as encoding $N$ elements one at a time.
@@
 The distribution is not what is being encoded.
-The variability is the information needed to go from the distribution, which is given, to an individual element.
+The variability is quantified by the information needed to go from the distribution, which is given, to an individual element.
 It is really the information gap from the population to an element.
-That is why some people say the entropy is ``lack of information", which is justified because, in a way, it is the information about the elements that the distribution cannot provide.
-But, again, this is just missing the point of what is being described.
-If one is not interested in identifying elements there is no ``missing information".
+That is why some people say the entropy is ``lack of information," which is justified because, in a way, it is the information about the elements that the distribution cannot provide.
+But, again, this is deceptive: if one is not interested in identifying elements there is no ``missing information."
 
@@
 
-This should suggest that the success of Shannon variability and its use in both information theory and physics does not warrant a fundamental change of perspective in what constitutes a physical object, as some physicists have claimed.
-It is true that any physical process can be used to process information, when properly encoded.
+The use of information in physics, then, does not warrant a fundamental change of perspective in what constitutes a physical object, as some physicists have claimed.
+It is true that any physical process can be used to process information, \emph{when properly encoded}.
 It is also true that scientific theories, in the end, are models that can capture only the aspects of nature that can be tested experimentally, the information extracted by the experiment, under suitable circumstances.
-Therefore the claim that information plays an essential role in physical theories has a valid basis.[cite Wheeler, Landauer] But it is also true that that data requires the context to be able to be understood: we need to know what is the subject of our experiment, how to prepare it and how to collect the data.
-The art of experimental science is neither contained in the mathematical description nor in the data collected.
+Therefore the claim that information plays an essential role in physical theories has a valid basis.\footnote{Comments like ``It is wrong, moreover, to regard this or that  physical quantity as sitting \emph{out there} with this or that numerical value'', ``the information thus solicited [by the experiment] makes physics and comes in bits'' by Wheeler\cite{Wheeler} or ``I am proposing that the ultimate form of the implementable laws of physics requires only operations available (in principle) in our actual universe'' by Landauer\cite{Landauer} go in this direction} But it is also true that that data requires the context in order to be understood: we need to know what the subject of our experiment is, how to prepare it and how to collect the data.
+The art of experimental science is contained neither in the mathematical description nor in the data collected.
 As information, in the information theoretic sense, requires that context to become intelligible, it cannot play a primary role.
-Therefore the claim that the universe itself is information does not follow.[cite Vedral, Davies]
-
-%Cite Vlatko Vedral: Decoding Reality: The Universe as Quantum Information. Information is "the fundamental building block of the Universe".
-
-%Cite Paul Davies Information and the Nature of Reality. Information "occupies the ontological basement"
-
-%Cite Wheeler It from bit.
-"It is wrong, moreover, to regard this or that  physical quantity as sitting "out there" with this or that numerical value it is absorbed." "The information thus solicited make physics and comes in bits"
+Therefore the claim that the universe itself is information\footnote{Comments like ``[Information is] the fundamental building block of the Universe'' by Vedral\cite{Vedral} or ``[Information] occupies the ontological basement'' by Davies\cite{Davies} go in this direction.} does not follow.
 
@@
 How does it work in that case?
-Wouldn't we need infinitely many bits to identify a value from a continuous distribution?
+Would we not need infinitely many bits to identify a value from a continuous distribution?
 
-\section{Continuous distributions}\label{sec_cont}
+\section{Continuous Distributions}\label{sec_cont}
 
-A standard quick and dirty way to extend distributions from discrete variables to continuous variables is to substitute weights with densities and the sums with integrals.
+A standard quick and dirty way to extend distributions from discrete variables to continuous variables is to substitute weights with densities and sums with integrals.
 Therefore, instead of having a discrete normalized distribution $\sum_i p_i = 1$, we have a continuous normalized distribution $\int \rho(x)dx=1$.
-While simply changing $- \sum_i p_i \log p_i$ to $- \int \rho(x) \log \rho(x) dx$ works in most cases, it is easy to get confused because the two expressions have significantly different properties.
+While simply changing $- \sum_i p_i \log p_i$ to $- \int \rho(x) \log \rho(x) dx$ works in most cases, leaving it at that misses a crucial point: the two expressions have significantly different properties.
 
@@
 A distribution over half a unit interval will correspond to minus one bit, since we would need to ``lose'' one bit of information to widen the variability back to one unit.
-In other words, the variability is measured by the number of bits needed to identify an element up to one unit.
+In other words, the variability is measured by the number of bits needed to identify an element up to one unit, up to the resolution (i.e.
+the granularity) given by that unit.
 
 This means that infinitesimally narrow distributions, like delta functions, do not work well with the Shannon variability as they would give minus infinite entropy.
-If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which have a very small but still finite support (i.e.
-region where the function is non-zero).
-We will assume all our continuous distributions are like that.
+If one wants entropy over continuous variables to be bounded, then one must work with continuous functions which  maybe have very small but still finite support (i.e.~region where the function is non-zero), which is what we will assume.
 
-This brings up a second problem: what happens if we change unit? If we change the reference, we expect that the entropy will change accordingly. In fact:
+This brings up a second problem: what happens if we change units? If we change the reference, we expect that the entropy will change accordingly. In fact we find
 \begin{equation}
 \begin{aligned}
-H(\rho(y)) &= - \int \rho(y) \log \rho(y) dy \\
-&= - \int \rho(x) \left|\frac{\partial y}{\partial x}\right| \log \left(\rho(x)\left|\frac{\partial y}{\partial x}\right| \right) \left|\frac{\partial x}{\partial y}\right| dx \\
-&= - \int \rho(x)  \log \left(\rho(x)\left|\frac{\partial y}{\partial x}\right| \right)  dx \\
-&= - \int \rho(x)  \log \rho(x)  dx - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
-&= H(\rho(x)) - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx \\
+H(\rho(y)) &= H(\rho(x)) - \int \rho(x)  \log \left|\frac{\partial y}{\partial x}\right|  dx 
 \end{aligned}
@@
 
-In physics, there is one thing that makes the Shannon variability work: phase-space.
-What happens in phase space is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components.
+In physics, what makes the Shannon variability work is phase space.
+What happens is that, under coordinate transformation $\hat{q}^i = \hat{q}^i(q^j)$, the $d\hat{q}^i = \frac{\partial \hat{q}^i}{\partial q^j } dq^j$ vary like vector components while $d\hat{p}_i = \frac{\partial q^j}{\partial \hat{q}^i } dp_j$ vary like covector components.
 Therefore the areas $d\hat{q}^i d\hat{p}_i = dq^j dp_j$ remain the same.
-The Jacobian determiants are unitary and the Shannon variability is invariant.
+This has a couple of important consequences.
+First of all, (1) the volumes are preserved during the transformation, which means (2) the Jacobian is unitary and (3) the Shannon variability is invariant.
+In fact, these three properties are mathematically the same property: each one implies all others.
+Second, something similar happens to the areas for each independent degree of freedom, meaning that the Shannon variability associated to the marginal distributions is also invariant.
+These properties can be taken to be the defining characteristics of phase space: phase space has exactly such geometrical properties.\cite{AoP2020HamiltonianEntropy}
 
-The use of phase-space in statistical mechanics is of paramount importance: over this space both the density $\rho(q^i, p
+The use of phase space in statistical mechanics is of paramount importance: over this space both the density $\rho(q^i, p
 _i)$ and its variability $H(\rho(q^i, p
@@
 
-The importance of phase space is often attributed to Liouville's theorem, that states that areas are conserved under Hamiltonian evolution.
+The importance of phase space is often attributed to Liouville's theorem, which states that areas in phase space are conserved under Hamiltonian evolution.
 While this is true, their invariance under coordinate transformations is more important as it is what makes them physically meaningful in the first place.
-In fact, how could we have conservation over time if we could not compare it in an observer independent way?
-This means that we always have to consider distributions over the full phase-space with the appropriate conjugate coordinates if we want to use $-\int \rho \log \rho$.
-If we use distributions over position and velocities or over momentum only, the expression would change for those particular variables in those coordinate systems.
+This means that we always have to consider distributions over the full phase space with the appropriate conjugate coordinates if we want to use $-\int \rho \log \rho$.
+If we use distributions over position and velocities or over momentum only, the expression will change for those particular variables in those coordinate systems.
 This can be, again, a source of confusion.\cite{Dunkel}
 
-We conclude noting that in quantum mechanics the Shannon variability, in the form of the Von Neumann entropy, is coordinate invariant as well.
-That is, given a density matrix $\rho$ then $S_{VN} = - \textrm{tr}(\rho \log \rho)$ is independent of the basis in which it is calculated.
+We conclude noting that in quantum mechanics the Shannon variability, in the form of the von Neumann entropy, is coordinate invariant as well.
+That is, given a density matrix $\rho$ then $S_\textrm{\tiny{vN}} = - \textrm{tr}(\rho \log \rho)$ is independent of the basis in which it is calculated.
 
-Now that we have seen how the Shannon variability works in the continuous case and the importance of phase space, we are ready to see how it can be used in statistical mechanics and how it relates to the Boltzman, Gibbs and Von Neumann entropies.
+Now that we have seen how the Shannon variability works in the continuous case and the importance of phase space, we are ready to see how it can be used in statistical mechanics and how it relates to the Boltzmann, Gibbs and von Neumann entropies.
 
@@
 
-As we saw in the previous sections, Shannon variability is a very general concept that applies to any type of distribution, no matter what the problem domain.
-It does not require us to define states, processes, reversibility and therefore it is conceptually different and independent from the entropy in thermodynamics and statistical mechanics.
-Yet it can, and is, applied successfully in statistical mechanics even in systems far from equilibrium.[cite] How does this work?
-
-As statistical mechanics aims to describe the collective behavior of a large number of physical systems (i.e.
-particles), the Shannon variability can be applied to their distribution and will correspond to the entropy of the system.
-Yet, we have to understand that there are two distinct ways to apply the concept simply because there are two types of distributions one can be interested in.
+Statistical mechanics aims to describe the collective behavior of a large number of physical systems (i.e.~particles), therefore asking what variability is expressed by such a collection is a well posed question.
+Yet, we have to understand that there are two distinct ways to apply the concept simply because there are two types of distributions in which one can be interested.
 We may consider the state of the whole system at a given time, and be interested in the distribution of the particles over all possible particle states.
-This will lead to the Boltzman entropy.
-We may consider a statistical ensemble, which is a large collection of independent copies of the system found in different states, and consider the distribution of these different instances over the states of the whole system.
-This will lead to the Gibbs or the Von Neumann entropy depending on whether the system is classical or quantum.
-Both these types of distributions are used and are therefore of interest, so we will look at both and clarify the potential sources of confusion.
+This will lead to the Boltzmann entropy.
+We may instead consider a statistical ensemble, which is a large collection of independent copies of the system found in different states, and consider the distribution of these different instances over the states of the whole system.
+This will lead to the Gibbs or the von Neumann entropy depending on whether the system is classical or quantum.
+Both these types of distributions are used and are therefore of interest, so we will examine both.
 
 Suppose we have a large number $N$ of particles taken from a normalized distribution $\rho(q^i, p_i)$.
-The space is the six dimensional phase-space for a single particle, sometimes called $\mu$-space.
+The space is the six-dimensional phase space for a single particle, sometimes called $\mu$-space.
 As we assume $N$ large, we can think of $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$ as the number of particles that are within a region $\Sigma$ of phase space.
@@
 
-The Boltzmann constant $k_B$ should not distract us: its role is simply to allow us to measure temperature in an appropriate unit.
+% http://przyrbwn.icm.edu.pl/APP/PDF/120/a120z6p04.pdf
+
+The Boltzmann constant $k_B$ should not distract us: its role is simply to allow us to measure temperature in an appropriate unit.\footnote{In the latest revision of the International Metric System of Units SI,\cite{chyla2011evolution} the Boltzmann constant is one of the defining constants.
+This means its value is assigned, rather than measured, in a way that properly defines Kelvin, the unit of temperature.
+The universal constants are essentially used as fixed starting points to define our measurement scales.
+}
 If one defines $T = \frac{\partial U}{\partial S}$, and measures temperature in Kelvin and energy in Joules, the above relationship forces us to measure entropy in Joules/Kelvin.
@@
 However, this is not the only possible definition.
-Instead of using $T$ as a primary thermodynamic variable, we can use $\beta = \frac{1}{k_B T}$.\footnote{There are a few reasons why one can argue this is a better choice, though ultimately it is still a matter of choice.} In this case, one would define $\beta = \frac{\partial S}{\partial U}$, measure energy in Joules, $\beta$ in inverse Joules and entropy would be dimensionless.
-As one can do all the thermodynamic calculations using just $\beta$, the constant $k_B$ is really just set by the unit system.
+Instead of using $T$ as a primary thermodynamic variable, we can use $\beta = \frac{1}{k_B T}$.
+In this case, one would define $\beta = \frac{\partial S}{\partial U}$, measure energy in Joules, $\beta$ in inverse Joules and entropy would be dimensionless.
+As one can do all the thermodynamic calculations using just $\beta$, the constant $k_B$ is really just set by the unit system.\footnote{As \cite{chyla2011evolution} notes:
+	\begin{quote}
+		Temperature characterizes the average thermal energy of particles in a certain ensemble of particles in the state of thermodynamic equilibrium.
+Therefore, in principle, one could express this quantity in terms of the unit of energy; this is actually practiced in many fields of physics, especially in statistical physics, where it is quite common to use $\Theta = k_B T$ or $\beta = 1 / k_B T$  as a measure of thermal energy, instead of temperature $T$ measured in Kelvins.
+	\end{quote}
+}
 
-This setting, distributions over single particle phase space, is the one used to derive the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the distribution of particles for an ideal gas.
-It provides a clear picture for what happens in simple cases: the equilibrium distribution is the one where the variability of the states of the particles is maximal.
-That is, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles.
-The information needed to identify a particle within the distribution at a given precision will increase during an irreversible process.
-The equilibrium is a statistical equilibrium, particles are moving around, but for one particle that moves in one direction, there is another one that moves in the opposite and the overall distribution remains the same.\footnote{This corresponds to the insight developed by Boltzmann that led to the creation of statistical mechanics.}
+Distributions over single-particle phase space constitute the setting used when deriving the Maxwell-Boltzmann distribution $\rho(q^i, p_i) = \left(\frac{\beta m}{2\pi} \right)^{3/2}e^{-\beta \frac{p_ip^i}{2m}}$, which is the distribution of particles for an ideal gas.
+As it evolves towards equilibrium, particles will spread out as much as they can under the constraints given by the energy, volume and number of particles, increasing the variability until it is maximized.
+The equilibrium is a statistical equilibrium, particles are moving around, but for any particle that moves in one direction, there is another one that moves in the opposite and the overall distribution remains the same.\footnote{This corresponds to the original insights developed by Boltzmann.}
 
-This approach, though, will only work in the limit of a large number of indistinguishable particles that are independently distributed.
-If $N$ is not large, there may not be another particle that moves in the opposite direction.
-If there are correlations, the motions may not cancel out.
-More precisely, suppose we have a probability distribution $\hat{\rho}$ for $N$ particles.
+This approach, though, will only work in the limit of a fixed large number of indistinguishable particles that are independently distributed.
+More precisely, suppose we have a joint probability distribution $\hat{\rho}$ for $N$ particles.
 If they are independently distributed, then the joint probability $\hat{\rho}=\prod_{i}\rho_i$ is the product of the distribution for each particle.
@@
 If $N$ is large, the number of particles in one region $N_\Sigma$ is very close to the expectation value $N\int_\Sigma \rho(q^i, p_i)dq^idp_i$.
-If these assumptions are not met, we cannot break the joint distribution into single particle ones, the fluctuations may become relevant thus requiring a more general account.\footnote{In [Jaynes] it is shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.} Moreover, a change of the number of particles or particles of different species cannot be captured by a single normalized distribution over single particle states.
+If these assumptions are not met, we cannot break the joint distribution into single-particle ones, particle number and type may change, the fluctuations may become relevant thus requiring a more general account.\footnote{Jaynes\cite{Jaynes} has shown that a single distribution over $\mu$-space will not recover the correct experimental values for entropy.}
 
 The more general setting, then, is the following.
-The macrostate is a probability distribution over all possible complete descriptions of the system, the microstates.
-That is, we have a distribution over the $6N$ dimensional phase space of $N$ particles, sometimes referred to as $\Gamma$-space, where each point represents the position and momentum of $N$ particles.
+The macroscopic state, or macrostate, is a probability distribution over all possible complete descriptions of the system, or microstates.
+That is, we have a distribution over the $6N$-dimensional phase space of $N$ particles, sometimes referred to as $\Gamma$-space, where each point represents the position and momentum of $N$ particles.
 The Gibbs entropy is $S_G = -k_B \int \rho \log \rho \, dq^idp_i$, which corresponds to the Shannon variability of the microstate distribution for a given classical macrostate.
@@
 
-There are a couple of problems in this picture that we have to be aware of.
+There are a couple of issues in this picture.
 The first is that $\Gamma$-space automatically assumes that all particles are distinguishable.
-This leads to the widely known problem of overcounting which needs to be addressed in the standard way.[cite] The second problem is that, though the state of each particle is given by position and momentum, we should not think of them as literally pointlike.
-As we said, this would correspond to delta Dirac distributions over phase-space which have minus infinite Shannon variability.
-It is more appropriate, both mathematically and conceptually, to think of indistinguishable particles as identical peaked distributions, each contributing the same amount of Shannon entropy.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.} The position and momentum corresponds to the center of mass of the particle.
+This leads to the widely known problem of overcounting which needs to be addressed in the standard way.
+The second problem is that, though the state of each particle is given by position and momentum, we should not think of them as literally pointlike.
+As we said, this would correspond to delta Dirac distributions over phase space which have minus infinite Shannon variability.
+It is more appropriate, both mathematically and conceptually, to think of particles as identically peaked distributions, each characterized by the same amount of Shannon variability.\footnote{Setting to $-\log h$ the entropy corresponding to each degree of freedom of these peaked distributions is a natural way to incorporate the effects of the uncertainty principle.} The position and momentum correspond to the center of mass of the particle.
 
-We note that some authors choose to interpret the probability distribution not as representing the frequency over a large amount of independent trials, but as the knowledge one has about the system.
-This would make the entropy an observer dependent quantity, instead of an objective notion.\footnote{Jaynes called it the anthropomorphic nature of entropy.} The issue is that, experimentally, the fluctuations we observe are not observer dependent.
-We believe the confusion comes from the, correct, realization that the system by itself is not enough to determine the thermodynamic variables that are needed to reach the equilibrium.
-In fact, Jaynes points out:
+We note that some authors choose to interpret the probability distribution not as coming from repeated independent trials, but as the knowledge one has about the system.
+This would make the entropy a subjective notion: each observer would have a different credence distribution, regardless of whether it fits the data, and therefore a different Shannon entropy.
+A physical quantity, however, must be the same for everybody, and the associated fluctuations we experimentally observe are indeed objective.
+So what is going on?
+
+We believe the confusion comes from the, correct, realization that the same system, under different conditions, will be described with a different set of thermodynamic variables.
+\footnote{Jaynes\cite{Jaynes} points out:
 	\begin{quote}
@@
 It is clearly  meaningless to ask, ``What is the entropy of the crystal?''  unless we first specify the set of parameters which define its thermodynamic state.
-\end{quote}
-This is a particular case of what we stated in the previous sections: a distribution, and therefore its variability, is contingent upon some arbitrary choices.
-But, again, once those choices are made, the distribution is objective.
+\end{quote}}
 The choice of the system, in statistical mechanics and thermodynamics, is enough to determine the state space for the microstates, but not enough to determine the set of macrostates that correspond to equilibria.
 We have to specify the process and the constraints that that process puts on the system.
-That is, under a different choice of process and constraints the microstates will fluctuate in different ways since we have changed the dynamics of the system.
-The question, then, is whether the system plus the process plus the set of constraints are enough to determine the entropy, or if we also need to mix in human knowledge.
-What we find is that they are sufficient, so the entropy does not depend on human knowledge.
+Under a different choice of process and constraints the microstates will fluctuate in different ways since we have changed the dynamics of the system.
+This is what we stated at the beginning: a distribution, and therefore its variability, is always contingent upon some arbitrary choices.
+Since the Gibbs entropy is the variability of a microstate within the distribution identified by the macrostate, it is not a property of the single microstate, it is not a property of the system, but it is a property of the system within that specific process, of the macrostate.
+Some authors refer to this fact by saying that entropy is not objective \footnote{Jaynes\cite{Jaynes,Jaynes2} called it the anthropomorphic nature of entropy.}.
+
+The issue is that subjective, in the context of probability theory, refers to Bayesian probability, which really means subjective: what one believes to be true regardless of what experimental evidence there is.
+This is not at all the same concept as the one outlined before.\footnote{Note how Jaynes\cite{Jaynes2} always put ``subjective'' in double quotes.} The system plus the process (which identifies the set of constraints) determines the ensemble and therefore the entropy, regardless of whether an agent knows what process was used.
+Even if one wants to give a Bayesian account, then, one has to give it in terms of an agent that has full and exact knowledge of the system and the process.
+But this is hardly subjective.
+In our view, this is a case of unfortunate word choice.
+It is much better to simply state that the entropy is a property of the ensemble, of the distribution, and not of the system.
+And if one understands that the Gibbs entropy is the variability of the microstate within the ensemble, then it is clear that it cannot be a property of the microstate itself, but it is a property of the ensemble.
 
 The case of a quantum system is formally similar to the classical one.
-Instead of a distribution over the $N$ particle phase space, we have a distribution over the Hilbert space for the quantum system which is represented by a density matrix operator $\rho$.
-The Von Neumann entropy is given by $S_{VN} = - \textrm{tr}(\rho \log \rho)$.
-This, expanded in a basis, becomes $S_{VN} = - \sum_i \rho_i \log \rho_i$ or $S_{VN} = - \int \rho(x) \log \rho(x)dx$ depending on whether the spectrum is discrete or continuous.
-The Von Neumann entropy corresponds to the Shannon variability of the microstate distribution for a given quantum macrostate.
+Instead of a distribution over the $N$-particle phase space, we have a distribution over the Hilbert space for the quantum system which is represented by a density matrix operator $\rho$.
+The von Neumann entropy is given by $S_\textrm{\tiny{vN}} = - \textrm{tr}(\rho \log \rho)$.
+This, expanded in a basis, becomes $S_\textrm{\tiny{vN}} = - \sum_i \rho_i \log \rho_i$ or $S_\textrm{\tiny{vN}} = - \int \rho(x) \log \rho(x)dx$ depending on whether the spectrum is discrete or continuous.
+The von Neumann entropy corresponds to the Shannon variability of the microstate distribution for a given quantum macrostate.
 
 The different entropies in statistical mechanics, then, all have a tight link to the Shannon variability.
-The Boltzmann entropy corresponds to the variability of the state of a particle within a given microstate, provided that there are a large number of independent and indistinguishable particles.
+The Boltzmann entropy corresponds to the variability of the state of a particle within a given microstate, provided that there are a large fixed number of independently distributed and indistinguishable particles.
 The Gibbs entropy corresponds to the variability of a classical microstate as constrained by the macrostate.
-The Von Neumann entropy is similar but corresponds to the quantum case.
+The von Neumann entropy is similar but corresponds to the quantum case.
 The characterization we gave to the Shannon formula, then, is readily applicable to statistical mechanics in a natural way.
-We find it more precise than the common characterizations, such as disorder or lack of knowledge, and, once internalized, it leads to less confusion.
 
 \section{Conclusion}
+
+In this paper we have seen that:
+\begin{itemize}
+	\item the Shannon entropy measures the variability of the elements within a given distribution, giving it a crisp intuitive meaning that is general and applicable to all branches of science
+	\item the expression is not arbitrary, as it is the only linear indicator for such a concept
+	\item it measures the variability by quantifying the number of yes/no questions one must ask to identify an element within the distribution, which corresponds to the number of bits needed to transmit or store that information
+	\item when properly applied to statistical mechanics, the variability leads to the Boltzmann, Gibbs and von Neumann entropies.
+\end{itemize}
+The characterization we gave to the Shannon formula, then, is more precise than the common characterizations, such as disorder, information or lack of knowledge, and it should lead to less confusion.
+It clarifies that the Shannon variability is an independent concept from the entropy of thermodynamics and statistical mechanics, and a link can be recovered only if properly applied.
+We find that this approach, once internalized, gives greater intuitive insight and also maps more readily to the mathematical details.
+
+\section{Acknowledgments}
+We would like to thank Juniar Lucien for insights into the physics education literature.
+This work is in connection to Assumptions of Physics, a larger project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
 

### 2021-02-05 Articles/2019-Entropy/Source/thermodynamics.tex
Reviewed thermodynamics paper

@@
 Counting the possible evolutions plays a fundamental role, in much the same way that counting states plays a role in statistical mechanics.
-In a deterministic and reversible process, there is one to one correspondence between initial and final states: the evolutions cannot merge or diverge.
+In a deterministic and reversible process, there is a one-to-one correspondence between initial and final states: the evolutions cannot merge or diverge.
 If the process is deterministic but not reversible, however, the same final state will be reached from different initial states, and the evolutions will merge.
@@
 
-\section{States, processes and evolutions}
+The paper is organized in three sections.
+In the first we will present a simple framework to characterize processes in general terms.
+In the following section we will add the definitions and assumptions that are specific to thermodynamic processes and rederive the basic laws.
+In the last section we see how the concepts presented relate to the different branches of mechanics.
 
-The first question we want to answer is: how can we characterize processes in the most general terms?
+\section{States, processes and evolutions}\label{sec_process_entropy}
+
+Our first task is to characterize processes in the most general terms.
 We assume we have a system and we want to study its behavior in a given set of circumstances.
 If the state space $X$ of the system represents all possible configurations the system can be found in, ultimately we want to characterize how the state changes in time.
-As depicted in figure \ref{fig_single_evolution}, we have a copy of the state space at each moment in time, and we call an \textbf{evolution} a trajectory $x(t)$ that for each moment in time returns the state of the system.
+As depicted in figure \ref{fig_single_evolution}, we have a copy of the state space at each moment in time, and we call an \textbf{evolution} a trajectory $\lambda : T \to X$ that for each moment in time returns the state of the system.
 
@@
 
-Not all evolutions will be possible under all conditions.
-Therefore we choose to characterize a \textbf{process} simply as the set of allowed evolutions, as we can see in figure \ref{fig_process}.
-Namely, a process tells us what can and cannot happen in a given set of circumstances.
-We also assume we have a measure $\mu$ that allows us to determine the size of a set of evolutions, to ``count'' evolutions.\footnote{The precise conceptual and mathematical definition for $\mu$ is a source of a number of technical considerations which we (will) address in a separate work.
-While some issues, like unit definition, are important to the physics, most have to do with the proper handling of infinities or the conditions for existence and uniqueness of the measure.
-Overall, they would distract from the key insights.} A state $x_t$ at a time $t$ identifies a set of of compatible evolutions $A(x_t)$, those that pass through that state at that time.
-That is, $x(t) = x_t$.
+Since not all evolutions will be allowed under all conditions, we characterize a \textbf{process} as the set of possible evolutions $E$, as we can see in figure \ref{fig_process}.
+Namely, a process tells us what can and cannot happen to a system in a given set of circumstances.\footnote{For example, a damped harmonic oscillator will allow trajectories whose oscillations degrade over time in a specific fashion.} We also assume we have a measure $\mu$ that allows us to determine the size of a set of evolutions, to ``count'' evolutions.\footnote{The full mathematical characterization is more complex: it will be a preorder that induces a family of measures, as to properly address sets of infinite relative sizes.
+However, these details do not change the overall intuition and would only distract from the key insights.} Note that a state $x_t$ at a particular time identifies a set of compatible evolutions $A(x_t) = \{ \lambda \in E \, | \, \lambda(t) = x_t \}$, those that pass through that state at that time.
 Using $\mu$ we can give a size to that set, which we can write as $\mu(A(x_t))$ or $\mu(x_t)$ for short.
@@
 
-Note that the probability of transitioning from one state to another during a process is linked to the allowed evolutions and their count.
+Note that the probability of transitioning from one state to another during a process is linked to the possible evolutions and their count.
 In fact, calculating the probability of finding the state $x_{t+\Delta t}$ at $t+\Delta t$ given $x_t$ at $t$ means determining the fraction of evolutions that pass through $x_t$ that also pass through $x_{t+\Delta t}$.
@@
 
-Having answered the question of how we characterize processes, we can categorize them.
-For example, we say a process is \textbf{deterministic} if the state at one time is enough to determine the state at all future times.
+This very minimal characterization is enough for some basic definitions.
+We say a process is \textbf{deterministic} if the state at one time is enough to determine the state at all future times.
 We can see an example of such a process in figure \ref{fig_determinism}.
@@
 The count of evolutions per state over time can never decrease: $
-\mu(x(t)) \leq \mu(x(t + \Delta t))$.
+\mu(\lambda(t)) \leq \mu(\lambda(t + \Delta t))$.
 
 Deterministic processes can be characterized by a law of evolution.
-That is, we can write $x(t+\Delta t) = f(x(t), t)$ given a suitable function $f$ which may, in general, be time dependent.
-We can do this because $P(x(t+\Delta t) | x(t)) = 0$ for all but one future state.
-The law of evolution identifies that state.
-That is, $P(x(t+\Delta t) | x(t)) = 1$ if and only if $x(t+\Delta t) = f(x(t), t)$.
+That is, we can write $\lambda(t+\Delta t) = f(\lambda(t), t)$ given a suitable function $f$ which may, in general, be time dependent.
+We can do this because, if we fix $x_t$, $P(x_{t+\Delta t} | x_t) = 1$ for one and only one future state and the law of evolution identifies that state.
+Insofar we are interested in making predictions, we are implicitly restricting ourselves to deterministic processes.
 
@@
 Conversely, we say a process is \textbf{reversible} if the state at one time is enough to reconstruct the state at all past times.\footnote{Note that in some literature the ability to reconstruct the past is called retrodictability and does not coincide with thermodynamic reversibility (the ability to undo a change).
-This distinction appears in the context of continuous quantities and it is precisely one of those technical considerations we (will) address in a more technical work.
-At this point we want to make it clear that our notion of reversibility will indeed correctly recover thermodynamic reversibility.
-It will do so without having to assume the existence of a reverse process or a symmetry under time-reversal.} We can see an example of such a process in figure \ref{fig_reversibility}.
+This distinction appears in the context of continuous quantities and it is one of those technical considerations we will address in a more technical work.
+Our definition does recover thermodynamic reversibility, without having to assume the existence of a reverse process or a symmetry under time-reversal.} We can see an example of such a process in figure \ref{fig_reversibility}.
 In this case, if two different evolutions happen to reach the same state at a given time, they will also share all past states: evolutions can diverge but cannot merge.
 The count of evolutions per state over time can never increase: $
-\mu(x(t)) \geq \mu(x(t + \Delta t))$.
+\mu(\lambda(t)) \geq \mu(\lambda(t + \Delta t))$.
 
@@
 A process that is both deterministic and reversible, like the one in figure \ref{fig_detrev}, will never contain evolutions that split or merge.
-The count of evolutions per state over time will be a constant: $\mu(x(t)) = \mu(x(t + \Delta t))$.
+The count of evolutions per state over time will be a constant: $\mu(\lambda(t)) = \mu(\lambda(t + \Delta t))$.
 
@@
 To finish, consider a \textbf{deterministic processes with equilibria}, like the one in figure \ref{fig_with_equilibria}.
-An equilibrium for a process is a state that, once it is reached, it is never left.\footnote{We will see later that this characterization of equilibrium is simplistic, but it will suffice for now.} In a process with equilibria all evolutions will reach, at some point, an equilibrium.
+An equilibrium for a process is a state that, once it is reached, it is never left.\footnote{This characterization of equilibrium, though amply used in thermodynamic literature, is simplistic and technically does not quite work.
+However, it will suffice for the purpose of this paper as a full treatment would again distract from the main points.} In a process with equilibria all evolutions will reach, at some point, an equilibrium.
 In the out-of-equilibrium phase, evolutions merge until they reach equilibria and can no longer merge.
@@
 
-We already reached an important result: with only a handful of simple definitions we have found a quantity, $\mu(x(t))$, that can never decrease in a deterministic process, stays the same if the process is also reversible and is maximized at equilibrium.
+We already reached an important result: with only a handful of simple definitions we have found a quantity, $\mu(\lambda(t))$, that can never decrease in a deterministic process, stays the same if the process is also reversible and is maximized at equilibrium.
 We have done so with no assumption on the type of system, the type of forces acting on the system, the internal composition of the system or the internal dynamics of the system.
@@
 
-Suppose we now have two systems, with their respective state spaces $X$ and $Y$.
+Suppose we have two systems, with their respective state spaces $X$ and $Y$.
 The state space of the composite system is the set of all possible pairs $X \times Y$.
-A process defined on the composite system will come with a measure $\mu_{X \times Y}$ that enables us to count evolutions $\{x(t), y(t)\}$ of the composite system.
+A process defined on the composite system will come with a measure $\mu_{X \times Y}$ that enables us to count evolutions $\lambda(t) = \{\lambda_x(t), \lambda_y(t)\}$ of the composite system.
 In general, the evolution of one system will affect the evolution of the other system.
 This will not happen, however, if the systems are independent.
-That is, as shown in \ref{fig_independence}, for each possible evolution $x(t)$ of the first system and $y(t)$ of the second system, there will be a possible evolution $\{x(t), y(t)\}$ on the composite.
+That is, as shown in \ref{fig_independence}, for each possible evolution $\lambda_x(t)$ of the first system and $\lambda_y(t)$ of the second system, there will be a possible evolution $\{\lambda_x(t), \lambda_y(t)\}$ on the composite.
 In other words,
@@
 \end{equation}
-the measure factorizes. We say two systems are \textbf{independent} at a given time $t$ if the evolutions per state factorizes: $\mu_{X \times Y}(x(t), y(t)) = \mu_X (x(t)) \mu_Y (y(t))$.
+the measure factorizes.
+We say two systems are \textbf{independent} at a given time $t$ if the evolutions per state factorizes: $\mu_{X \times Y}(\lambda(t)) = \mu_X (\lambda_x(t)) \mu_Y (\lambda_y(t))$.
 
@@
 In addition, the process entropy will be additive for independent systems.\footnote{As before there are a number of technical details that need to be addressed, most of all the units for $\mu$ which will influence the 0 point for $S$.
-We will address this in the full framework.}
+We leave these details to future works.}
 
 Under very minimal conditions, we have found a quantity that has all the right properties for entropy and that is linked with determinism and irreversibility in a plain, intuitive and physically meaningful way.
-
-\section{System independence and state entropy}
-
-The careful reader will have noticed that the process entropy will depend on the process, not just on the state.
+There is one remaining issue: the process entropy will depend in general on the process, not just on the state.
 That is, the same description of the same system may be associated with more or fewer evolutions depending on the conditions.
@@
 
-If the system is a composite there are additional considerations.
+For composite systems there are additional considerations.
 Different processes may put different degrees of correlation between subsystems, therefore the same equilibria for the subsystems may correspond to different process entropy values for the composite.
@@
 
-We can then assign a unique process entropy to states of equilibria and to composite states of independent equilibria.
-Under these conditions we define the \textbf{state entropy} $S(x)$ which corresponds to the process entropy that the system reaches at equilibrium.
-But why would we want to assign a unique process entropy to states in the first place?
-Why would this be interesting?
-There are three main reasons.
+We can therefore assign a unique process entropy to states of equilibria and to composite states of independent equilibria.
+Under these conditions we define the \textbf{state entropy} $S : X \to \mathbb{R}$ which corresponds to the process entropy that the system reaches at equilibrium.
+Note that both equilibrium and independence lead to entropy maximization, so the state entropy can be thought as the maximal process entropy for the system.
+In all other cases, the process entropy will be lower meaning that that description for that system in those conditions gives us more information because of possible correlations within the system or with other systems.\footnote{This recovers the intuition of entropy as ``lack of information'' and that entropy maximization represents the case where knowledge of the system is fully represented by the given constraints.
+However, in our case this is not a subjective matter, rather an objective description of what the process may or may not allow.}
 
-The first is process combination.
-Suppose that we are studying a system and we change the circumstances surrounding it.
-It first evolves according to $A$ and at $t_0$ we switch to process $B$.
-If $x(t)$ is an evolution of the combined process $AB$, then for $t\leq t_0$ it must be an evolution for $A$ and for $t\geq t_0$ it must be an evolution for $B$.
-Moreover, the number of total evolutions must remain the same.
-Therefore the state $x(t_0)$, the state where we are joining the evolutions, must have a well defined entropy that is independent of the process, or not all the evolutions would match.
-If we want to mix and match from a set of possible processes, then the process entropy at the junction must be only a function of the state.
+This link between state entropy and independent equilibria opens a series of questions on the fundamental link between the two which go beyond the scope of this paper.
+However, we want at least give a partial answer to one: why is assigning a unique process entropy important?
+One answer is process combination.
+Suppose that we are studying a system and we change the circumstances surrounding it: it first evolves according to process $A$ and at $t_0$ it switches to process $B$.
+If $\lambda(t)$ is an evolution of the combined process $AB$, then for $t \leq t_0$ it must also be an evolution for $A$ and for $t \geq t_0$ it must also be an evolution for $B$.
+This means that at the junction state $x_{t_0} = \lambda(t_0)$ the count of evolution has to match: $\mu_{AB}(x_{t_0})=\mu_A(x_{t_0})=\mu_B(x_{t_0})$.
+If we want to mix and match from a set of possible processes, then the process entropy at the junction must be only a function of the state, independent of the process.
+As we'll see in section \ref{sec_mechanics}, these type of consideration are already embedded in the mathematical structure of the state spaces of classical mechanics, quantum mechanics, thermodynamics and statistical mechanics.
 
-The second reason is that, upon close inspection, measurements are never defined at an instant of time.
-Therefore $x(t)$, the state at an instant of time, is really a shorthand for $x[(t - dt/2, t + dt/2)]$, the state defined over a small open interval $dt$ around $t$.
-We can talk about a single state within that interval only if no state variables change and the system is fully characterized by those state variables alone.
-Therefore writing $x(t)$ implicitly assumes two things: that the system is an equilibrium of the processes that happen at a faster scale than our measurements of said system; that further description (e.g.
-the internal dynamics of the constituents at a finer resolution) is not needed to describe the process.
+\section{Thermodynamics}\label{sec_thermodynamics}
 
-The third reason is that by state we implicitly mean a description of a system and only about that system.
-If there are correlations between that system and others at a time, then the description of one will tell us more than that system alone.
-Proper states are then defined in the case where the system is independent, where all the correlations have been removed.
-This is true also for correlations with the environment and correlations between subsystems.
-But the condition of independence is precisely the condition that maximizes entropy, making it uniquely defined.
+The above discussion sets the stage for our derivation of thermodynamics.
+We saw how state entropy is the maximized process entropy, a quantity that can never decrease under deterministic processes.
+Our task now is to extend those concepts in a way that recovers thermodynamics.
 
-All three insights point toward the same direction: the very notion of state seems to be intrinsically associated to the notion of equilibrium and state entropy.
-This would suggest that all states for all systems are, in a sense, equilibria.
-But what exactly is this sense?
-What does it mean that a mechanical system, classical or quantum, is at equilibrium?
+We can motivate this extension by considering process composition of composite systems.
+The idea is that we want to describe the situation in which a composite\footnote{Mathematically, systems under composition form a Boolean algebra.
+The join operation $X \cupdot Y$ returns the composite; the meet operation $X \capdot Y$ returns the smallest system contained by both; the lattice ordering $X \subseteq Y$ tells us whether $X$ is a part of $Y$ and so on.} system $X \cupdot Y$ is subject to process $A$ until time $t_0$, after which each subsystem evolves independently: $X$ according to process $B$ and $Y$ according to process $C$.
+At $t_0$, then, we not only have to reconcile the count of evolutions for the composite, but for each subsystem as well.
+Therefore $X$ and $Y$ must be independent and in equilibrium at time $t_0$.
+In our framework, thermodynamic equilibrium handles precisely this situation.
+The remaining ingredient is the notion of extensive quantities, which again serves to characterize part-whole relationships.
 
-Even in thermodynamics, equilibrium is usually an ill-defined concept.
-It is typically associated with the one coming from dynamical systems (i.e.
-the system no longer changes), but this doesn't quite work.
-In fact, it is not even relativistically consistent.
-If we have a volume of gas in thermodynamic equilibrium, we would like to say it is in equilibrium for any observer, not just the one in its rest frame.
-A boosted observer, then, would see an equilibrium where the position is indeed changing.
-In the presence of gravitational forces, this motion would not even be linear.
-The notion of equilibrium as ``nothing is changing'', then, is not quite right and should be amended.
-But how?
-The previous discussion would hint toward system independence, but why should system independence imply an equilibrium?
+We define a \textbf{thermodynamic process} a particular type of deterministic process with equilibria that satisfies two additional requirements:
+\begin{itemize}
+	\item all parts of the systems are in equilibria and are independent from each other
+	\item there is a non-trivial dependence between the initial state of one part and the final state of another (i.e.
+there is a way to change the initial state of one part such that the final state of another is effectively changed).
+\end{itemize}
+The first requirement tells us that process entropy is maximized at all levels of the system.
+The second requirements serves to rule out the case where the systems are isolated from each other.
+We say a system is in \textbf{thermodynamic equilibrium} if it is part of the output of a thermodynamic process.
+We say two systems are in thermodynamic equilibrium with each other if they are parts of the same system in thermodynamic equilibrium.\footnote{Mathematically, the set of all systems in the same thermodynamic equilibrium is a principal ideal $X\!\!
+\downarrow$.}
 
-The insight is that we can define something as a system only when we are able to take it apart from the rest and study it independently.
-Therefore we need, at least conceptually, a process that takes that something and makes it independent from the rest, decouples it, makes it autonomous.
-Our ability to define a system implicitly assumes the existence of such a ``decoupling'' process.
-Note that if the system is already independent, the decoupling process does nothing: in this sense independent systems are an equilibrium of a decoupling process.
+To get a feel for the definition, let us rederive the fact that thermodynamic equilibrium is transitive, also known as the zeroth law.
+Suppose system $X$ is in thermodynamic equilibrium with system $Y$ and system $Y$ is in thermodynamic equilibrium with system $Z$.
+Then $X$ and $Y$ are parts of the output of the same thermodynamic process and so are $Y$ and $Z$.
+This means that $X$ and $Z$ are parts of the output of the same thermodynamic process, and therefore are in thermodynamic equilibrium.
 
-This ``decoupling'' process is what maximizes entropy.
-If the system is truly rendered independent, then we cannot know more than the description given by the state at that particular time.
-We cannot know how the system came to be in that state, we cannot know anything more about the subsystems, other systems or the environment than what the state of the system strictly tell us.
-When the system is independent, the states tell us the least possible information about everything else and the evolution of anything else.
-
-The definition of a particular system, then, is conditional on the existence of particular processes: those that render the system well defined.
-We can talk about the motion of a ball because we are on earth at 20 Celsius.
-If we were on the surface of the sun, we would not be able to talk about the ball in the first place.
-To define a system requires us to define a boundary, which should remain the same during the process.
-In mathematical terms, the state space must be a symmetry (i.e.
-an invariant, an equilibrium) of a group (i.e.
-a set of transformations, a set of processes).
-
-The full mathematical development of these concepts is left for another work, but the conceptual foundations are simple and direct.
-Grouping quantities into cohesive units that ``belong together'', into systems, comes with a set of practical requirements: we need a way to prepare and study them independently.
-This notion of ``independence equilibrium'' is what we will use as a foundation for ``thermodynamic equilibrium'', and is generalized to all systems.
-It is the necessary and sufficient condition to assign state entropies.
-In this light, entropy maximization is not a universal law of the universe, but a necessary requirement to build models.
-We can only study the state of objects that allow process entropy maximization.
-What remains is to answer one last fundamental question: is this notion of state entropy at all related to thermodynamic entropy?
-
-\section{Thermodynamics}
-
-The notion of equilibrium as independence is in general not sufficient to recover thermodynamic equilibrium. So, how are they related? When does the second arise?
-
-We saw how the notion of state as independence equilibrium came naturally when considering process composition of a single system.
-When the process changes, we need to reconcile the count of evolution at the junction.
-We now want to consider the scenario of process composition of composite systems.
-The idea is that we want to describe the situation in which two systems come together, evolve as a composite for some time, and then proceed to evolve separately.
-In this case, at each juncture we have to reconcile the evolutions of both the whole and each of the parts.
-We need to have an independence equilibrium at both levels.
-What we'll find is that thermodynamics studies this case.
-But let's proceed in order.
-
-If our goal is to characterize part-whole relationships, a crucial question is: how are the state variables of the whole related to the state variables of the parts?
-We define a \textbf{thermodynamic system} one that follows these rules:
+To characterize part-whole relationships between states, we focus on particular types of systems, which we call \textbf{thermodynamic systems}, that satisfy these additional requirements: 
 \begin{itemize}
-	\item states (i.e.
-independence equilibria) are identified by a set of extensive quantities (i.e.
-additive under system composition)\footnote{It may be possible that the existence of extensive quantities can be derived from the composability of thermodynamic equilibria.
-Note, in fact, that we typically construct measurement scales of quantities (such as energy, mass, volume, distance, ...) by preparing unit samples and then adding or dividing them.
-Assuming the existence of thermodynamic equilibria may enable us similar constructions without extra assumptions.
-(XXX: any work related to this?) Nevertheless, for this work, we consider the existence of extensive quantities an extra assumption.
+	\item the state space is formed by states of equilibria
+	\item states are identified by a set of extensive quantities (i.e. additive under system composition)
+	\item one such quantity, called internal energy, is conserved under all thermodynamic processes.\footnote{It may be possible that some or all of these properties can be derived from the composability of thermodynamic equilibria.
+Nevertheless, we consider them additional assumptions for this work.
 	}
-	\item one such quantity is called internal energy
-	\item internal energy will be conserved under any deterministic process for that system.
 \end{itemize}
@@
 
-For a thermodynamic system, then, the state entropy can be expressed as $S(U, x^i)$ a function of the extensive quantities.
-This is simply the equation of state in entropic form, which is the starting point of Gibbs formulation of thermodynamics.
-[CITE] We can therefore define the relationships
-\begin{align}
-\beta &= \frac{\partial S}{\partial U} = \frac{1}{k_B T} \;\;\;\;\;\; - \beta X_i = \frac{\partial S}{\partial x^i}
-\end{align}
-where $k_B$ is Boltzmann's constant. We find:
-\begin{align}
-dS &= \frac{\partial S}{\partial U} dU + \frac{\partial S}{\partial x^i} dx^i \\
-&= \beta dU - \beta X_i dx^i \\
-dU &= k_B T dS + X_i dx^i
-\end{align}
-which are the usual thermodynamic relationships, except that, for us, the state entropy $S$ is dimensionless.
-
-As we need to characterize states of both whole and parts, we need to characterize the part-whole relationship of the equilibrium.
-We say that a thermodynamic system $A$ in \textbf{thermodynamic equilibrium} has the following properties:
-\begin{itemize}
-	\item the system $A$ is in an independence equilbrium
-	\item all parts of $A$ are in thermodynamic equilibrium
-	\item if another composite system $B$ is in thermodynamic equilibrium and $A$ and $B$ share a part, then the composite of $A$ and $B$ is in thermodynamic equilibrium.
-\end{itemize}
-We say two or more systems are in thermodynamic equilibrium with each other if their composite is in thermodynamic equilibrium.\footnote{To understand the mathematical structure a bit better, the set of all systems $S$ is a Boolean algebra.
-Element $0$ represents ``nothing'' while element $1$ represents ``everything''.
-The join $\OR$ operation is system composition while the meet $\AND$ returns the smallest system contained by the two.
-Therefore $A \OR 1 = 1$: any system composed with everything gives us everything.
-As a lattice, the partial order $\leq$ is system inclusion.
-A thermodynamic equilibrium essentially identifies a topmost system, with all the subsystem and only the subsystems as part of the same equilibrium.
-That is, for every thermodynamic equilibrium there exists a system $A$ such that every $B \leq A$ is in the same equilibrium while every $C \nleq A$ is not in the same equilibrium.} A \textbf{thermodynamic process} is a process with equilibria where the final state is a thermodynamic equilibrium.
-We stress that we are not saying that two systems, in general, always allow for thermodynamic equilibria.
-However, we contend that these are the type of processes we can mix and match to create composite processes of composite systems.
-What we want to explore is what happens if we had such systems and processes, at least to a first approximation.
-
-To get a feel for the definition, let us rederive the fact that thermodynamic equilibrium is transitive, also known as the zeroth law.
-Suppose system $A$ is in thermodynamic equilibrium with system $B$ and system $B$ is in thermodynamic equilibrium with system $C$.
-Then the composites $A \times B$ and $B \times C$ are respectively in thermodynamic equilibrium.
-Since they share a part, $B$, this means that $A \times B \times C$ is in thermodynamic equilibrium.
-But $A \times C$ is a part of $A \times B \times C$, so it is in thermodynamic equilibrium.
-Which means $A$ is in thermodynamic equilibrium with $C$.
-
-The characterization of thermodynamic equilibrium we give is manifestly a part-whole relationship.
-It extends the notion of independence equilibrium to a composite system so that the equilibrium of the whole and the equilibrium of its parts are essentially the same equilibrium.
-We find that clarifying this kind of part-whole relationships makes for a good conceptual starting point not only for thermodynamics, but also for the other branches of physics.
-This should not come as a surprise as reductionism (i.e.
-explaining the whole through its parts) plays an important role in science.
-
-%Under these assumptions/definitions, we can recover the laws of thermodynamics. Let $A$ and $B$ be two thermodynamic system in thermodynamic equilibrium and let $AB$ be the composite. We have:
-%\begin{equation}
-%\begin{aligned}
-%U_{AB} &= U_A + U_B \\
-%S_{AB}(U_{AB}) &= S_A(U_A) + S_B(U_B) \\
-%&= \sup \{S_A(x) + S_B(y) \, | \, x+y= U_{AB}\} \\
-%\frac{1}{kT_A} = \beta_A = \frac{\partial S_A}{\partial U_A} &= \frac{\partial S_B}{\partial U_B} = \beta_B = \frac{1}{kT_B}
-%\end{aligned}
-%\end{equation}
-%which follows the standard textbooks derivations and assumes the new equilibrium is not on a boundary point.
-In this case, equilibrium is achieved if and only if $T_A$ and $T_B$ are equal.
-Since equilibria are unique, this will also tell us that all equations of states have to be concave in $U$.
-If $B$ is in thermal equilibrium with another system $C$ we also have $T_B$ is equal to $T_C$ and therefore $A$ and $C$ are also in thermal equilibrium.
-This recovers the zeroth law.
-%
-%There is, however, a problem.
-We assumed the equilibrium is not a boundary point.
-It could be that system $A$ reaches its lowest energy state at a finite $T_A$ while $B$ can go to lower temperatures.
-This would mean that $B$
-
-A natural question, based on the previous definitions, is: to what extent can thermodynamic processes change the energy and the entropy of each system?
+The next step is to characterize to what extent can thermodynamic processes change the energy and the entropy of each system.
 Intuitively, there are two ``orthogonal'' directions for a change of energy: constant entropy and maximal entropy change.
@@
 
-We say a thermodynamic system $R$ is a \textbf{thermal reservoir} if its equation of state can be sufficiently approximated by $S = \beta U = \frac{U}{k_B T}$.
-Internal energy is the only state variable and the only way the entropy of the system can change.
-We call \textbf{heat} the energy exchanged by a thermal reservoir and we note $Q = - \Delta U$ as the energy lost by the reservoir during a thermodynamic process.
+We say a thermodynamic system $R$ is a \textbf{thermal reservoir} if the internal energy $U_R$ is the only state variable.
+In this case, the equation of state is simply $S_R = S_R(U_R)$ and $\Delta S_R = \int \beta_R dU_R$.\footnote{The existence of unique equilibria between reservoirs forces $S_R(U_R)$ to be strictly monotonic.} We call \textbf{heat} $Q = - \Delta U_R$ the energy lost by the reservoir during a thermodynamic process.
 
 We say $M$ is a \textbf{mechanical system} if the state entropy is a constant for all states.
-For a thermodynamic system that is also mechanical, we therefore have  $S(U, x^i) = k_M$ for all its states.
+For a thermodynamic system that is also mechanical, we therefore have  $S_M(U_M, x^i_M) = k_M$ for all its states.
 A mechanical system cannot change entropy and all states can be connected through a deterministic and reversible process.
@@
 \begin{aligned}
-dU &= k_B T dS + X_i dx^i = X_i dx^i.
+dU_M &= k_B T_M dS + X_{Mi} dx^i_M = X_{Mi} dx^i_M.
 \end{aligned}
@@
 The energy of the mechanical system is effectively \emph{stored} in one of the other state variables and can be later retrieved.
-We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U$ as the energy gained by the system during a thermodynamic process.\footnote{Note that $\frac{\partial S}{\partial U} = 0 = \beta = \frac{1}{k_B T}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite.
+We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U_M$ as the energy gained by the system during a thermodynamic process.\footnote{Note that $\frac{\partial S_M}{\partial U_M} = 0 = \beta = \frac{1}{k_B T_M}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite.
 While this ``temperature'' has little to do with thermometers, the fact that we can always transform all work into heat but not vice-versa is just a specific case of being able to move energy from higher to lower temperature.}
 
-We can study how changes of energy and entropy are related to each other.
-Consider a system composed of a generic system $A$, a thermal reservoir $R$ and a mechanical system $M$.
-In this setting, work represents the energy exchanged at constant entropy while heat represents exchanged energy associated with entropy change.
+With these definitions, we can study how changes of energy and entropy are related to each other.
+Consider a system composed of a generic system $X$, a thermal reservoir $R$ and a mechanical system $M$.
 
@@
 \begin{aligned}
-&\Delta U_A + \Delta U_R + \Delta U_M = 0 \\
-&\Delta U_A - Q + W = 0 \\
-&\Delta U_A = Q - W.
+&\Delta U_X + \Delta U_R + \Delta U_M = 0 \\
+&\Delta U_X - Q + W = 0 \\
+&\Delta U_X = Q - W.
 \end{aligned}
 \end{equation}
-This recovers the first law of thermodynamics.
+In this setting, work represents the energy exchanged at constant entropy while heat represents exchanged energy associated with entropy change. This recovers the first law of thermodynamics.
 
@@
 \begin{aligned}
-&\Delta S_A + \Delta S_R + \Delta S_M \geq 0 \\
-&\Delta S_A \geq - \Delta S_R \\
-&\Delta S_A \geq \beta_R \Delta Q \\
-&\Delta S_A \geq \frac{\Delta Q}{k_B T_R}.
+&\Delta S_X + \Delta S_R + \Delta S_M \geq 0 \\
+&\Delta S_X \geq - \Delta S_R \\
+&\Delta S_X \geq \int \beta_R (- dU_R) \\
+&\Delta S_X \geq \int \beta_R d Q \\
+&k_B \Delta S_X \geq \int \frac{d Q}{T_R}.
 \end{aligned}
@@
 The third law has many equivalent formulations.
-[CITE] One of them states that the entropy of any system has a non-negative lower bound which is reached by the lowest energy state.
+[CITE] One states that the entropy of any system has a non-negative lower bound which is reached by the lowest energy state.
 If the ground state is non-degenerate, such as a perfect crystal, its entropy is equal to zero.
-We can more easily recover it in this form.
+With some work, we can recover this form.
 
@@
 Now we show that either all thermodynamic systems must have a lowest energy state or they must have a highest energy state.
-Suppose, in fact, we had two systems: $A$ has no maximum energy while $B$ has no minimum energy.
-Since $S_A(U_A)$ is concave, it must be a monotonically increasing function of $U_A$.
-This means that $\beta_A$ is always positive.
-Conversely, $S_B(U_B)$ must be monotonically decreasing, which means $\beta_B$ is always negative.
-If we put the two systems in contact, the energy of $B$ will keep decreasing and the energy of $A$ will keep increasing and no equilibrium will ever be reached.
+Suppose, in fact, we had two systems: $X$ has no maximum energy while $Y$ has no minimum energy.
+Since $S_X(U_X)$ is concave, it must be a monotonically increasing function of $U_X$.
+This means that $\beta_X$ is always positive.
+Conversely, $S_X(U_X)$ must be monotonically decreasing, which means $\beta_X$ is always negative.
+If we put the two systems in contact, the energy of $Y$ will keep decreasing and the energy of $X$ will keep increasing and no equilibrium will ever be reached.
 Therefore all thermodynamic systems either have a minimum or a maximum energy.
-At this point the difference is a sign convention, so we can assume all thermodynamic systems have a minimum energy.
+At this point the difference is a sign convention, so we can assume by convention that all thermodynamic systems have a minimum energy.
 
 To finish, we note that a perfect crystalline structure fits the case where the system has only one possible evolution, so it is a zero entropy state.
-If there is degeneracy, there are different ways the equilibrium can be realized, instead, there are different evolutions that will correspond to the same equilibrium which means the entropy will be greater than zero.
+If there is degeneracy, there are different ways the equilibrium can be realized and there are different evolutions that will correspond to the same equilibrium, which means the entropy will be greater than zero.
 This recovers the third law.
 
-The laws of thermodynamics have been recovered, and we have done so only by characterizing how equilibria and state variables behave under system composition.
-Apart from that, we have made no assumptions on what type of system, the type of forces acting on the system, the internal composition of the system or the internal dynamics of the system.
-No reference to statistical mechanics or quantum mechanics is needed: the theory stands on its own.
+We have thus recovered the laws of thermodynamics, and we have done so only by characterizing how equilibria and state variables behave under system composition.
+Apart from that, we have made no assumptions on the type of system, the type of forces acting on the system, the internal composition of the system or the internal dynamics of the system.
+No reference to classical mechanics, quantum mechanics or statistical mechanics is needed: the theory stands on its own.
 This fits the original spirit of thermodynamics, which was to find a set of laws that would be independent fn the ultimate dynamics of the constituents.
+This also explains the success of these ideas in fields outside physics.
 Additionally, the concepts we used are directly related to the physical explanation, and the laws are recovered immediately, not after pages and pages of calculations.
-The combination of these features makes this approach directly useful, for example, in an introductory physics course: a coherent self-contained explanation can be given in terms of simple definitions.
+The combination of these features makes this approach directly useful, for example, in an introductory  course: a coherent self-contained explanation can be given in terms of simple definitions.
 
-% \subsubsection{XXX Null states and thermodynamic zero}
+\section{Connections to mechanics}\label{sec_mechanics}
 
-% * Use null state as thermodynamic zero.
-% * Must have zero entropy.
-% * All extensive quantities for the null state must be zero.
-% * Extensive quantities can be zero for a propert state (i.e. number of particle of a particular species).
-% * Intensive quantities are derivative: they also have an absolute zero.
-% * All extensive quantities must be non-negative?
+We now want to return to the link between evolution count, state entropy and state spaces.
+In particular, we want to show how this link exists not only in thermodynamics, but also in classical mechanics, quantum mechanics and statistical mechanics.
+This is important not only to show how process entropy relates to the two main ways to define entropy in statistical mechanics, but to give a sense that all geometrical and probabilistic structure assigned to state spaces are ultimately capturing evolution counts in one way or another.
+This is important for our bigger project, Assumptions of Physics, as we want to create a common framework for all physical theories.
 
-% Sample text
+We previously defined mechanical systems as those for which all states have the same entropy.
+While one does not usually discuss the entropy of single classical or quantum states, it is easy to see that that must be the case.
+In quantum mechanics, for example, the von Neumann entropy of every pure state is zero.
+In classical mechanics, a uniform measure is used to count states, which means all states must contribute the same entropy.
+Additionally, in both theories, given two states, we can always (at least mathematically) find a deterministic and reversible process (Hamiltonian evolution and unitary evolution are entropy preserving) that starts with the first and ends in the second.
+So the idea that mechanical systems are associated with a constant entropy, though seldom explicit, is already embedded in the established theories.
 
-%Zero entropy is usually associated to a non-degenerate ground state and the lower bound on energy is usually thought as a consequence of quantum mechanics.
-We saw how, in our framework, these requirements stem from the definitions.
-In fact, we can repeat the argument and find that \emph{all} extensive quantities.
-We also note that, while one typically focuses on temperature and entropy having an absolute scale, other thermodynamics variable, such as volume and pressure, have a well defined zero.
-Is there a better way to frame these issues?
-We believe there is.
+The geometrical structure of classical mechanics, the symplectic structure of phase space, can be understood to be exactly the structure we need to count evolutions when the evolution is deterministic and reversible and the degrees of freedom are independent.
+In that case state entropy and process entropy coincide, but not in general.
+In dissipative processes, like a damped harmonic oscillator, this relationship breaks down.
+As the system loses energy, the evolutions gradually concentrate around the attractors.
+In terms of count of states, areas shrink over time, so counting states gives us a decreasing number, which is at odds with the idea of entropy increasing when progressing towards an equilibrium.
+What happens is that as the system evolves, each region will correspond to the same number of evolutions, which will now be spread over fewer states: the process entropy (i.e.
+the logarithm of the evolutions per state) increases as evolutions become more concentrated.
 
-%System composition $\times$ is a closed operation (i.e.
-given any two systems returns another system) that is commutative (i.e.
-$A \times B = B \times A$) and associative (i.e.
-$A \times (B \times C) = (A \times B) \times C$).
-Mathematically it makes sense to ask whether it has an identity (i.e.
-$A \times 0 = 0 \times A = A$), which we note as $0$ since the operation is commutative (i.e.
-additive).
-In fact, while the state space of the composite is the product of the subsystems, system composition in terms of systems is better thought of as an addition: we are putting things together.
-The identity, then, is naturally the ``null'' system: the empty system.
-We can keep adding nothing, and the system is unchanged.
-This makes system composition a commutative monoid, but not a group since there is no ``inverse'' system: the one that combined with the original gives us nothing.\footnote{Note that particles/anti-particle annihilation leaves something: the energy.
-So the ``anti''-system is not the ``inverse'' system.}
+The mathematical structure of quantum mechanics also can be understood in terms of evolution counts.
+The inner product of quantum mechanics gives us the conditional probabilities of a ``black box'' process (i.e.
+the projection/measurements) that, given an initial state, gives us a final one.
+As we saw, conditional probabilities are ratios of evolution counts: what fractions of the evolutions that start in one state end in another.
+In other words, the inner product tells us how the evolutions coming in and coming out of these processes are to be ``joined'' with the preceding and following ones.
+Moreover, note that these processes are projections, meaning they give the same result if applied more than once.
+Eigenstates of projections, then, are equilibria of the process the projection represents.
+Or, in other words, equilibria are symmetries of the process.
 
-%The null system has only one possible configuration, which we call the null state, so it can only have one possible evolution.
-The null state of a null system is inherently in equilibrium and the state entropy is zero.
-In fact, the values for all extensive quantities must be zero since composing the null system with another null system still gives us a null system.
-Therefore, for example, $U_0 + U_0 = U_0$.
+In statistical mechanics, one uses the above structures to define the entropy of an ensemble.
+There are currently two ways to do this.
+The first, commonly referred to as the fundamental postulate of statistical mechanics, is to take the logarithm of the state count of the region of phase space identified by the constraints (e.g.
+the energy of the system is within a particular bound).
+We can relate this to the process entropy in the following way.
+Suppose that the constraints are such that each microstate either is compatible or incompatible with the macrostate.
+Then the count of evolutions of the macrostate is the sum of all the evolutions of its microstates.
+Now suppose that, when the constraints are defined, the system is at equilibrium and independent.
+Then the process entropy is equal to the state entropy.
+Since the system is mechanical, the state entropy is the same for all states.
+This means that, if the conditions above are met, the process entropy is, up to a constant, the logarithm of the state count.\footnote{If the system starts independent and at equilibrium, for example, the entropy of a final equilibrium would correspond to the logarithm of the size of the basin of attraction: each state corresponds to a single evolution that ends in the attractor.}
 
-% The intensive quantities are defined as derivatives, which means their zero is also absolute: is the case where changing the extensive quantity leads to no change in entropy.
-This means that, in thermodynamic, all quantities must be absolute, not only temperature or entropy is absolute, but this tells us that all thermodynamic quantities must be absolute.
-Volume, pressure, number of particles, we already think of them as having a well defined zero.
+The second way to define entropy is through the Gibbs and von Neumann entropies, which essentially are the application of the Shannon entropy in the classical and quantum case.
+This can be related to the process entropy in the following way.
+Suppose we have a macroscopic equilibrium which can be understood as a dynamical equilibrium of the microscopic description.
+That is, the macroscopic equilibrium is a set of evolutions whose microscopic description can be characterized by a stable probability distribution $\rho(x)$.
+An evolution $\lambda(t)$ for the microstate, then, is an dense sequence of infinitely many microstates whose recurrence matches $\rho(x)$.
+Since the state entropy is the maximal process entropy, we are counting all possible evolutions which means all possible permutations.
+The Shannon entropy can be understood as the count of all possible permutations of an infinite sequence, of all possible evolutions.
 
-%Some confusion comes from energy.
-In mechanics energy in the form of the Hamiltonian does not have an absolute zero, an arbitrary constant can be added with no ill effect.
-But rescaling internal energy by a constant $\hat{U}$ would make it no longer an extensive quantities: $(U_A + \hat{U}) + (U_B + \hat{U}) = U_{AB} + 2 \hat{U} \neq U_{AB} + \hat{U}$.\footnote{XXX: someone must have already said this} Now the question is whether
+The above insights show that the framework we are setting up is not just a nicer retelling of thermodynamics, but it is something more fundamental that underlies all physical theories.
+Most, if not all, experiments are described as a sequence of different steps (e.g.
+one or more preparation, manipulation and measurement) stitched together.
+Making sure that the evolution count matches at every junction is a very simple and necessary bookkeeping requirement, and therefore it makes sense it plays such a major role in defining state spaces.
 
-\section{Connections to mechanics}
+Therefore it makes sense that countMaking sure that the evolution count  Stitching together those phases into a single process
 
-While recovering thermodynamics from straightforward physical definitions has its own merits, we believe the approach can be used to offer new insights into classical mechanics, quantum mechanics and statistical mechanics.
-So we want to conclude with a few remarks on the subject that give a feel for how we intend to extend the work in the fully developed framework.
+The idea of using states as junctures when processes change is at the core of scientific experimentation since most experiments are
 
-We have argued the states of a system are implicitly independence equilibria and have an associated state entropy.
-One does not typically speak in these terms for classical and quantum states.
-Is this warranted and useful?
-Do all mechanical systems match our requirement to have constant entropy?
-If one looks closely, these facts are already implicitly assumed.
-For quantum mechanics, in fact, the von Neumann entropy of every pure state is zero.
-In classical mechanics, a uniform measure is used to count the accessible states; additionally the Gibbs entropy is invariant under translations in phase space.
-This means that all states must contribute the same entropy.
-Additionally, in both theories, given two states, we can always (at least mathematically) find a deterministic and reversible process (i.e.
-a Hamiltonian evolution or a unitary evolution) that starts with the first and ends in the second.
-So the idea that mechanical systems are associated with a constant entropy, though seldom explicit, can already be deduced from the currently established theories.
-
-Another natural question is: how does the process entropy relate to the fundamental postulate of statistical mechanics, that entropy is the logarithm of the count of states?
-Suppose we give the description of a system at a given time (e.g.
-the energy of the system is within a particular bound).
-This will identify a set of possible evolutions and therefore a value for the process entropy.
-Also, this is the sum (or integral) of the process entropy associated with each state not ruled out by that description: every evolution will have to go through one of those states at that time.
-Now suppose the system is, at that time, independent from all others.
-Then the process entropy for each state is the state entropy, so the process entropy for that description is the sum (or integral) of the state entropy.
-Additionally, suppose the system is mechanical, then the process entropy for the description is, up to a constant, the count of states.
-Entropy as the count of microstates is easily recovered.
-Yet, it is recovered under specific conditions.
-
-These conditions can be met in different scenarios.
-We could have an independent mechanical system under a deterministic and reversible process.
-Then counting evolutions is the same as counting states, up to a constant, at any moment in time.
-In another scenario, the process is not deterministic and reversible but is isentropic, meaning that the process entropy is constant in time.
-If we have an independent mechanical system, its evolution will necessarily be isentropic since, whatever state we evolve to, the state entropy will always be the same.
-Now suppose the states compatible with our description mix with each other and only with each other.
-Each evolution will change state in time, but it will never go outside of the set, so the description works both as a constraint and a description of the dynamic equilibrium.
-At each time, the same description will correspond to the same states, so the process entropy will be constantly given by the count of states.
-
-Then one may ask: how are these two entropies different?
-Imagine that the mechanical system starts independent but the dynamics is non-reversible leading the dynamical system to an equilibrium, as in a damped harmonic oscillator.
-The process entropy at the initial state will coincide with the count of states, but not at later times.
-As the evolutions concentrate around the equilibrium, the number of states becomes fewer so the number of evolutions per state increases.
-The entropy of the equilibrium, assuming that it is ever reached, is essentially the size of the basin of attraction at the initial time.
-The entropy around the equilibrium, then, increases in time which is consistent with the process being irreversible.
-Conversely, note that the areas in phase space shrink as the available states become fewer and fewer, which would mean entropy as the log of the state count decreases under an irreversible process.
-It was actually this paradox that pushed us into considering evolutions as more fundamental than states and, retrospectively, it is the obvious choice.
-
-This example also shows how critical the notion of independence is to define state entropy.
-The evolution of a damped harmonic oscillator is not independent from the evolution of the systems that are absorbing the dissipated energy, which are in fact correlated: the higher the energy of the oscillator, the higher the energy lost.
-The same description at different times (e.g.
-the system has reached equilibrium) does not provide exactly the same information on the system or other systems.
-In these cases, the state entropy does not allow us to compare apples to apples while the process entropy always does.
-
-Having seen how the process entropy relates to the count of states, one may naturally ask: how does it compare with the Gibbs/Shannon entropy?
-Suppose you cannot distinguish the dynamics of a system under a given time resolution.
-That is, in principle the system has a well defined trajectory $x(t)$ but you can only measure quantities of the form $A[x(t)] = \int_{t_0}^{t_0 + \Delta t} a(x) dt$ (e.g.
-average energy, average number of particles, ...).
-A permutation of states within the trajectory will not affect any of the measurement outcomes.
-Therefore the best we can do, given enough of those measurements, is to measure the frequency $\rho(x)$ each state is visited within the time resolution.
-Now we ask, given a specific distribution $\rho(x)$, how many possible evolutions $x(t)$ would that correspond to?
-That means counting the number of permutations $W[\rho]$ we can construct.
-The process entropy will be $\log W$.
-The Shannon/Gibbs entropy is the limit of that expression if we take $x(t)$ to be continuous.
-
-The approach, then, is not simply a nicer retelling of thermodynamics, but it can help us better understand the fundamental connection between all these theories.
+ maostate preparation)can help us better understand the fundamental connection between all these theories.
 We stress that we are not arguing that all other notions of entropy should be abandoned in favor of process entropy: quite the contrary.
@@
 In this paper we gave a conceptual overview for a new approach to recover thermodynamics based on the simple idea of counting evolutions, the possible ways a system can evolve under a process.
-We saw that the logarithm of evolution count under a deterministic process has all the properties we associate with entropy and can serve as a foundation for thermodynamics but may also be further developed into a common foundation for the other physical theories.
+We saw that the logarithm of evolution count under a deterministic process has all the properties we associate with entropy.
+We have seen that the evolution count is maximized at equilibrium and for independent system and only in those cases it depends only on the state.
+A well defined evolution count is necessary if we want to be able to compose processes, since the incoming evolutions needs to match the outgoing evolutions for all states at the junction.
+Thermodynamics studies exactly the case of composite processes of composite systems under some simplifying conditions.
+This gives a simple way to rederive thermodynamics in a straight-forward and conceptually insightful way.
 
-Our aim is to fully develop a formal framework based on these ideas, which will give us a general theory for states and processes.
-Conceptually, we start with all that can be observed at all moments of time; organize the possible descriptions into systems/degrees of freedom by grouping those that ``go together'' in the sense that they can be studied, at least under some circumstances, independently from the rest; define state spaces as ``templates'' of those descriptions that can be instantiated over time.
-This will force us to spell out the hidden assumptions one always has when defining systems, states and processes.
-The goal is to then recover the different physical theories with additional assumptions on the system or on the processes being studied, as we have done here for thermodynamics.
-
-Mathematically, we will build on our work on the logic of experimental verifiability, that already gives us the basic building blocks (i.e.
-topologies, $\sigma$-algebras).
-This needs to be extended to allow us to count possibilities in general and evolutions in particular.
-Technically, we need a unified tool that generalizes notions of measure theory, geometry, probability and information theory as these cover specific aspects that do not work in the general case.
-
-It is important to us that these additional efforts, while making the mathematics more precise, will not obscure the physical motivation and intuitive understanding of the framework.
-These additional efforts will not change the nature of the insights presented.
+Moreover, we have seen that the state spaces of classical, quantum and statistical mechanics include geometrical and probabilistic structures that can all be understood in terms of evolution count.
+Our hope and long term goal is to develop a formal theory based on these ideas that identifies a minimum set of requirements for the definition of a state space.
+This theory can be further refined with additional assumptions in different ways to recover the different physical theories.
+We believe this would help us better understand the right questions to ask in within the foundations of physics.
 

### 2021-02-07 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Adding notes

@@
 For this reason, $\hat{\rho} $ remains coordinate-dependent.
+[XXX: more than locally....
+remove locally]
 %Fortunately, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
@@
 Both follow from imposing the motion equivalence condition.
-In equation 5, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (which is connected to the gauge-redundancy of the conjugate momentum).
+In equation 5, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (which is connected to the gauge-redundancy of the conjugate momentum) [XXX: and the magnetic force -- ].
 Just as Curiel does, we then used this relationship to derive a quadratic form for the Hamiltonian (Equation 7).
@@
 These include models of a classical photon.
-In general, Hamiltonian systems outside the hyper-regular domain correspond to cases where we cannot express the density of states in terms of position and velocity.
+[XXX: a massless particle that moves at constant speed] In general, Hamiltonian systems outside the hyper-regular domain correspond to cases where we cannot express the density of states in terms of position and velocity.
 These are cases where the weak equivalence condition (equation~\ref{weak_equivalence}) fails.
@@
 His definition takes satisfying certain kinematical constraints as fundamental.
-In contrast, we privilege a system having a unique solution, i.e.
+[XXX: state defined at equal time, velocity is not] In contrast, we privilege a system having a unique solution, i.e.
 being deterministic and reversible.
@@
 In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy the proper kinematical constraints (e.g.
-$v^i = \dot(x)^i$).
+$v^i = \dot{x}^i$).
 Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will allegedly still satisfy these kinematical constraints \parencites*[311]{Curiel}.
@@
 Yet, it is difficult to understand what a Lagrangian system outside the hyper-regular domain could mean.
-For outside this domain, the Euler--Lagrange equations do not have a single, unique solution.
+[XXX: typically, non-regular lagrangian are broken into regions where they give one solution] For outside this domain, the Euler--Lagrange equations do not have a single, unique solution.
 As an example, consider the case where the Lagrangian $L$ equals zero.
-This Lagrangian corresponds to a Hamiltonian that also equals zero, entailing that $dH = 0 $.
+This Lagrangian gives rise to a Hamiltonian that also equals zero, entailing that $H = 0 $.
 By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.
@@
 Both kinds of systems are Hamiltonian thanks to deterministic and reversible evolution.} Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections.
+
+[XXX: missing bit about changing assumption to quantum]
+
+Why we prefer this view:
+\begin{itemize}
+	\item Explains why statistical mechanics is best done on phase space and the role of the symplectic form in counting states for each degree of freedom (two for the price of one)
+	\item Clarifies the physical role of each mathematical element (no hiding in mathematical jargon)
+	\item Allows for a more geometrical explanation of the action principle
+	\item Gives a simple understanding of gauge transformations
+	\item Change of one assumption recovers quantum mechanics (clarifies the boundary between the two) refer to long paper
+\end{itemize}
 

### 2021-02-07 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
edits to all sections but 5.1 (coordinate-dependence) and Curiel section [before merging with GC edits from Monday]

--- 2021-02-02 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 
-Philosophers have recently shown that we can interpret Lagrangian and Hamiltonian mechanics as theoretically equivalent in a precise mathematical sense.
+The intellectual richness of classical mechanics manifests itself in an abundance of formulations: with Newtonian, Hamiltonian, Lagrangian, and Hamilton--Jacobi being the most prominent.
+Butterfield \parencites*[]{Butterfield} has aptly referred to these formulations as different \textit{frameworks} for treating classical systems.
+Recent work in philosophy of science has focused largely on how these formulations relate to one another, with the aim of arriving at a deeper understanding of classical mechanics.
+By considering the physical and conceptual differences between these formulations, we can better distinguish physical features from artifacts of our representations.
+
+In the context of comparing Hamiltonian with Lagrangian mechanics, three interpretive stances have emerged.
+\textcites[]{North} has argued for privileging Hamiltonian mechanics, on the grounds that it requires less mathematical structure to formulate.
+In contrast, \textcites[]{Curiel} recommends privileging Lagrangian mechanics, arguing that it isolates physically-significant kinematical constraints that all classical systems must satisfy.
+Finally, we can interpret Lagrangian and Hamiltonian mechanics as theoretically equivalent in a precise mathematical sense.
 Within a subclass of classical mechanical systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh}{Barrett2}.
 Categorical equivalence provides a natural interpretive criterion for theoretical equivalence because it yields mutual inter-translatability between theory formulations.
-At first glance, this equivalence motivates the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain.
-Yet, it allows for physical and conceptual differences to arise outside the hyper-regular domain.
-%Once identified, these differences can even affect how we interpret hyper-regular systems.
-
-Within this wider domain, \textcites[]{North} has argued for privileging Hamiltonian mechanics, on the grounds that it requires less mathematical structure to formulate.
-In contrast, \textcites[]{Curiel} recommends privileging Lagrangian mechanics, arguing that it isolates physically-significant kinematical constraints that all classical systems must satisfy.
-Here, we will argue that there are indeed physical grounds for privileging Hamiltonian mechanics over Lagrangian mechanics, their categorical equivalence notwithstanding.
-Whereas North's argument relies on a problematic criterion for comparing structure---based on the size of symmetry groups \parencites[]{Swanson}---our argument avoids this pitfall.
+This equivalence motivates the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain.
+% It is interesting to think how the other formulations fit in, e.g. Hamilton--Jacobi. Is there a categorical equivalence result for this formulation in some domain?
+
+Here, we will argue that there are indeed physical grounds for privileging Hamiltonian mechanics over Lagrangian mechanics, their categorical equivalence notwithstanding.\footnote{Barrett shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics are categorically equivalent provided that we define categories of models over the tangent and cotangent bundle formulations, respectively \parencites*[1181-82]{Barrett2}.
+However, he also shows that if we consider a more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are \textit{not} categorically equivalent \parencites*[1182-83]{Barrett2}.
+Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation, thereby avoiding trivialization worries from categorical equivalence.} Whereas North's argument relies on a problematic criterion for comparing structure---based on the size of symmetry groups \parencites[]{Swanson}---our argument avoids this pitfall.
 We show that a system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian.
@@
 Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this difference has interpretive consequences even within the hyper-regular domain.
-Furthermore, the construction at the heart of this argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density, along with a novel interpretation of the principle of stationary action within the Hamiltonian formalism.
-
-Understanding the relationship between Hamiltonian and Lagrangian mechanics depends partly on how one defines ``classical systems." Different definitions can result in different interpretive relationships.
+The construction at the heart of our argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density.
+We show that the invariant density sits more naturally within the Hamiltonian framework, providing another reason to privilege it over Lagrangian mechanics.
+% Our formalism also provides a novel interpretation of the principle of stationary action within Hamiltonian mechanics.
+
+Section~\ref{privileging} provides an informal overview of our argument for privileging Hamiltonian mechanics.
+We introduce three physical conditions that define classical systems while distinguishing Hamiltonian from Lagrangian mechanics.
+Section~\ref{fundamentality} explains how our position differs from North's.
+Specifically, our argument does not require metaphysical commitments to fundamental structure or perfectly natural properties.
+Section~\ref{derivation} describes in detail the derivation of Hamiltonian mechanics for systems that satisfy two physical conditions: infinitesimal reducibility and deterministic/reversible evolution.
+We then show (Section~\ref{motion}) that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
+This completes our primary argument for privileging Hamiltonian mechanics.
+
+Subsequent sections provide additional reasons to privilege Hamiltonian mechanics, along with illustrating further interpretive benefits of our approach.
+Section~\ref{density} explains how the invariant phase space density plays a special role in the Hamiltonian framework.
+The density allows us to compare and track the amount of material in different regions of phase space.
+In contrast, Lagrangian mechanics does not support these important properties of the density.
+Despite being coordinate-invariant, the density itself is coordinate-dependent.
+We therefore argue that some quantities are physically significant despite being coordinate-dependent.
+Finally, Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework.
+We provide a geometric interpretation of the action principle, which applies even outside the hyper-regular domain.
+We also consider and rebut objections stemming from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
+Along the way, we consider both the gauge redundancy of the conjugate momentum (Section~\ref{fundamental}) and of the Lagrangian (Section~\ref{gauge}).
+
+% We define classical systems in terms of basic physical conditions that they satisfy. These conditions justify the appropriateness of certain mathematical structures.
+
+% Understanding the relationship between Hamiltonian and Lagrangian mechanics depends partly on how one defines ``classical systems." Different definitions can result in different interpretive relationships.
 For instance, restricting classical systems to the hyper-regular domain precludes certain interpretive considerations.
@@
 
-Section~\ref{deterministic} introduces the joint condition of deterministic and reversible evolution.
+% Section~\ref{deterministic} introduces the joint condition of deterministic and reversible evolution.
 We will show that a classical system which satisfies this condition is Hamiltonian: its phase space is a symplectic manifold that evolves according to Hamiltonian evolution  (i.e.
@@
 
-We begin in Section~\ref{privileging} with an informal overview of our argument for privileging Hamiltonian mechanics.
-We introduce the key physical conditions that distinguish Hamiltonian from Lagrangian mechanics.
-Section~\ref{fundamentality} explains how our position differs from North's.
-Specifically, our argument does not require commitments to fundamental structure or perfectly natural properties.
-Section~\ref{derivation} describes in detail the derivation of Hamiltonian mechanics for systems that satisfy two physical conditions: infinitesimal reducibility and deterministic/reversible evolution.
-We then show (Section~\ref{motion}) that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
-Section~\ref{density} explains how the invariant phase space density plays a special role in the Hamiltonian framework.
-In contrast, we show that in configuration space, the invariance of the density is coordinate-dependent (arising from the fact that the Jacobian equals unity in Cartesian coordinates but not in general).
-Finally, Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework.
-We provide a geometric interpretation of the action principle, which applies even outside the hyper-regular domain.
-We also consider and rebut objections stemming from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
-% In configuration space, it is in some sense an accident that the density is invariant
-
 %Note that the photon does not belong to the hyper-regular domain. Furthermore, if you are not in the hyper-regular domain, then there is no longer a one-to-one map between position and velocity.
@@
 
-We now outline the three main steps of our argument, summarized in Figure~\ref{diagram}.
+Our argument has three main steps, summarized in Figure~\ref{diagram}.
 The left column lists three basic physical conditions: infinitesimal reducibility, determinism/reversibility, and motion equivalence.
@@
 
-First, a system is \textit{infinitesimally reducible} provided that the states of its constituent, infinitesimal parts completely specify the overall state of the system.
+First, a system is \textit{infinitesimally reducible} provided that the state of the whole system is equivalent to specifying the states of its parts (and vice versa).
+Infinitesimal reducibility distinguishes classical from quantum systems, where the latter are irreducible (we cannot specify the whole state by specifying the states of its subsystems).
+In the widest sense, a classical system is any infinitesimally reducible system.
 Colloquially, we can say that the whole state is equal to the sum of its parts, where we take these infinitesimal parts to be classical particles.
-When this condition is satisfied, we can define a distribution $\rho$ over particle states.
+
+If a system is infinitesimally reducible, then we can define a distribution $\rho$ over particle states.
 Furthermore, since particle states are independent of a choice of units, this distribution must be invariant.
@@
 The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
-[To do: Clarify the scope of our treatment: is it restricted to $N$-particle classical mechanics?
-Or does it also handle rigid bodies and classical continual mechanics?
+
+[To do: Clarify the scope of our treatment: at least in this paper, it seems to be restricted to $n$-particle classical mechanics.
+Does it also handle rigid bodies and classical continuum mechanics?
 Cite Mark Wilson 2013 on these issues].
@@
 This condition entails that the invariant distribution $\rho$ is in fact a conserved phase space density.
-We can thus model such systems using Hamiltonian mechanics, defined on a phase space with a symplectomorphism.
-These two conditions thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution.
+We will show that a classical system which satisfies this condition is Hamiltonian: its phase space is a symplectic manifold that evolves according to Hamiltonian evolution  (i.e.
+according to a symplectomorphism).
+We thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution.
 Section~\ref{deterministic} justifies this correspondence.
-[To do: on the topic of failures of determinism, cite Earman  \textit{a primer on determinism}, Baez singularities paper, or Norton's dome paper].
 
@@
 As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
-
-Importantly, some systems fail to satisfy the condition of motion equivalence.
-It is not always true that we can determine the evolution of a system in phase space from its trajectory in physical space.
-This is because knowing a system's position and velocity is not in general sufficient for knowing its momentum.
-For instance, knowledge of the physical trajectory of a photon, when treated as a particle, is insufficient for knowledge of its phase space evolution, since we cannot determine the momentum of a photon from only its trajectory through space.
-In general, different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
+% % Would be ideal if we can provide some literature support for this notion of motion equivalence. E.g. connecting it with a discussion in Marsden and Abraham foundations of mechanics.
+
+We cannot always determine the evolution of a system in phase space from its trajectory in physical space, since some systems fail to satisfy the condition of motion equivalence.
+In general, knowing a system's position and velocity does not suffice for knowing its momentum.
+Considering a photon as a classical particle, we cannot determine its momentum from its trajectory through space.
+Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
+Different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
 In contrast, the trajectories in phase space (encoded by the Hamiltonian vector field) uniquely define the Hamiltonian up to a constant.} Hence, the set of classical systems that satisfy motion equivalence is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics.
 For at least this reason, Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
-[To do: verify that the scenario of two different Lagrangians---giving rise to the same trajectories in physical space with different energies---arises only when the motion equivalence condition fails.
+% [To do: verify that the scenario of two different Lagrangians---giving rise to the same trajectories in physical space with different energies---arises ONLY when the motion equivalence condition fails.
 See Curiel 2014, page 42 for his take on this kind of scenario.]
+
+Of course, understanding the relationship between Hamiltonian and Lagrangian mechanics partly depends on how one defines ``classical systems." Different definitions can result in different interpretive relationships.
+On our account, Newtonian systems are the widest domain of classical systems, including any system that is infinitesimally reducible.
+Hamiltonian mechanics picks out the subset of Newtonian systems that follow deterministic and reversible evolution.
+Our treatment thereby sets aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \parencites*[]{Earman,Baez,Norton}.} Additionally, we set aside dissipative systems, i.e.
+those that fail to conserve energy.
+Possessing a deterministic and reversible evolution is equivalent to being non-dissipative, i.e.
+conserving energy.
+Consequently, our derivation of Lagrangian mechanics in Section~\ref{motion} takes place within the context of conservative systems, so we do not consider systems that satisfy inhomogenous Euler--Lagrange equations.
+Although the Lagrangian formulation applies to dissipative systems \parencites[]{Smith,Curiel}, this is not what we will mean by ``Lagrangian systems."\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime.
+We will view non-dissipative systems as Newtonian, but neither Lagrangian nor Hamiltonian.
+Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
+%To avoid verbal disputes, it is essential to clarify what falls within the scope of ``classical systems." One way to delimit classical systems is to specify a set of equations that they satisfy---such as Hamilton's equations or the Euler--Lagrange equations.
+We will instead follow a different approach by delimiting classical systems based on the physical conditions that they satisfy.
+
+ %[To do: on the topic of failures of determinism, cite Earman  \textit{a primer on determinism}, Baez singularities paper, or Norton's dome paper].
 
@@
 
-% Overall summary: A classical system is an infinitesimally reducible system.
+% %Overall summary: A classical system is an infinitesimally reducible system.
 The state space of these systems can be modeled by a symplectic manifold (the physical assumption of infinitesimal reducibility corresponds to the mathematical structure of symplectic manifolds).
@@
 
-\subsection{(Non-metaphysical) fundamentality and structure}
+\section{Fundamentality, \textit{sans} metaphysics}
 \label{fundamentality}
 
-% % To do: slip in a citation to Dorr/Hawthorne 2014, to note the independent metaphysical challenges that the notion of naturalness faces, and likewise for a metaphysical notion of fundamentality.
+% % done; slip in a citation to Dorr/Hawthorne 2014, to note the independent metaphysical challenges that the notion of naturalness faces, and likewise for a metaphysical notion of fundamentality.
 
@@
 As indicated at the start of Section~\ref{privileging}, we use ``fundamentality" in a non-metaphysical fashion.
-On our usage, differences in fundamentality amount to differences in generality.\footnote{Of course, we allow for other non-metaphysical uses of ``fundamental," tracking differences in epistemic or conceptual primacy.
+On our usage, differences in fundamentality amount to differences in generality.\footnote{We also embrace other non-metaphysical uses of ``fundamental," tracking differences in epistemic or conceptual primacy.
 For instance, we can think of the canonical position and momentum observables as ``epistemically fundamental" because they suffice to completely specify a classical state, fixing its values for all other observables.
-\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{fundamental}, discusses further senses in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires that fewer physical conditions are satisfied.
+\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{fundamental} discusses further senses in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
 Compared with North's \parencites*[]{North} account of structure, our definition has two distinct advantages.
@@
 
-Methodologically (but not metaphysically), our approach has many parallels to North's.
-We agree with North that it is fruitful to identify both invariants and---more generally---``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}.
+Despite our different commitments, our approach has many methodological parallels to North's.
+We agree with North that it is fruitful to identify both invariants and ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}.
 Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North}.
@@
 According to North, ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" \parencites*[66]{North}.
-These notions of fundamentality and intrinsicality correspond to differences in joint-carving, with more fundamental or intrinsic features being more joint-carving, or in Lewis's \parencites*[]{Lewis1983} terminology, more natural.
+These notions of fundamentality and intrinsicality correspond to differences in joint-carving, where more fundamental or intrinsic features are more joint-carving (or in Lewis's \parencites*[]{Lewis1983} terminology, more natural).
 
@@
 Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration.
-As we will see, precisely this scenario occurs in the context of Lagrangian mechanics.
-Hence, whereas North views Lagrangian mechanics as containing ``excess, superfluous structure" \parencites*[75]{North}, we take this additional structure as unnecessary but still physically significant.
-It corresponds to the physical assumption of motion equivalence, which applies to many classical systems.
+Precisely this scenario occurs in the context of Lagrangian mechanics, which requires the additional condition of motion equivalence to hold.
+Whereas North views Lagrangian mechanics as containing ``excess, superfluous structure" \parencites*[75]{North}, we take this additional condition as unnecessary but still physically significant.
+Thus, while it is illuminating to identify the minimum mathematical structures necessary for a formulation, additional structure is not automatically rendered physically insignificant.
+%It corresponds to the physical assumption of motion equivalence, which applies to many classical systems.
 %This characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system. 
@@
 
-Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as a formal device for encoding physical information or as a calculational tool.} Whether Hamiltonian phase space structure is metaphysically fundamental does not matter for the questions we seek to answer.
+Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as a formal device for encoding physical information or as a calculational tool.} Whether Hamiltonian phase space structure is metaphysically fundamental (in a classical world) does not matter for the questions we seek to answer.
 Regardless of its metaphysical status, Hamiltonian mechanics is physically more fundamental, and we will argue for its epistemic and conceptual primacy as well.
-Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references].
+Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [to do: mention these different Hamiltonian formalisms and provide references].
 Each of these frameworks has different pragmatic costs and benefits.
-Hence, it is plausible that there may be no single, unique necessary mathematical structure for mechanics, but rather a family of related structures.
-There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions.
-Furthermore, each of these formalisms references some ``unphysical" coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.\footnote{Tulczyjew's reformulation of Hamiltonian mechanics supplies a further reason for viewing symplectic geometry---and hence Hamiltonian mechanics---as conceptually fundamental.
-According to Teh and Tsementzis, this reformulation shows that  ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh}.}
+Hence, it is plausible that there may be no single, uniquely necessary mathematical structure for mechanics, but rather a family of related structures.
+In other words, there may be no canonical minimal structure for formalizing the relevant physical models and physical assumptions.\footnote{Tulczyjew's reformulation of Hamiltonian mechanics supplies a further reason for viewing symplectic geometry---and hence Hamiltonian mechanics---as conceptually fundamental.
+According to Teh and Tsementzis, this reformulation shows that  ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh}.}  Additionally, each of these formalisms references some ``unphysical" coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.
+We return to the physical significance of certain coordinate choices in Section~\ref{fundamental}.
 %The mathematical structure that you ``need" might not be unique.
@@
 North also introduces a technical notion of comparative structure, based on the symmetry group of statespace.
-According to this criterion, statespace A has more structure than statespace B if the symmetry group of A is a proper subgroup of B's symmetry group \parencites[87-88]{North}.\footnote{This statement follows the most charitable reconstruction of North's criterion given by \textcites[]{Swanson}.
-North sometimes speaks in terms of comparing the dimensions of symmetry groups, which they note leads to trivial counterexamples.} A smaller symmetry group entails that it is in some sense ``harder" to preserve the space's structure, indicating that it has more mathematical structure to be preserved.
-These symmetry groups are defined in terms of the coordinate transformations on the space that preserve geometric structure, amounting to re-coordinatizations (redescriptions) of that structure.
+According to this criterion, statespace A has more structure than statespace B if the symmetry group of A is a proper subgroup of B's symmetry group \parencites[87-88]{North}.\footnote{This criterion follows from the most charitable reconstruction of North's argument, given by \textcites[]{Swanson}.
+North sometimes speaks in terms of comparing the dimensions of symmetry groups, which they note leads to counterexamples.} A symmetry group that is a proper subgroup entails that strictly fewer transformations preserve the space's structure.
+This indicates that statespace A has more mathematical structure to be preserved.
+These symmetry groups are defined in terms of the coordinate transformations that preserve the space's geometric structure, amounting to re-coordinatizations (redescriptions) of that structure.
 For Lagrangian systems, North takes the relevant symmetry group to be that of the point transformations, and for Hamiltonian systems, she assumes that canonical transformations play this role.
@@
 As they note, Galilean boosts are often symmetry transformations of Lagrangian systems (in the sense of preserving the equations of motion), but they are not point transformations.
-Additionally, the two-body (Kepler) problem admits canonical transformations that map physically distinct solutions to each other (specifically, solutions whose orbits possess the same energy but different eccentricity).
-Hence, we cannot straightforwardly compare the structure of Lagrangian and Hamiltonian mechanics in terms of the symmetry groups of their statespaces.
+Additionally, the two-body (Kepler) problem admits canonical transformations that map physically distinct solutions to each other (specifically, solutions whose orbits possess the same energy but different eccentricity).\footnote{For more details on the classical two-body problem, see \textcites[]{Belot}.} Hence, we cannot straightforwardly compare the structure of Lagrangian and Hamiltonian mechanics in terms of the symmetry groups of their statespaces.
 
 Advantageously, our criterion for comparing Hamiltonian and Lagrangian mechanics avoids the criticisms that Swanson and Halvorson develop against North.
-For the sense in which Hamiltonian mechanics has less structure than Lagrangian mechanics is simply one of logical strength: a system must satisfy strictly fewer physical conditions in order to be Hamiltonian.
+Hamiltonian mechanics has less structure than Lagrangian mechanics simply in the sense of logical strength: a system must satisfy strictly fewer physical conditions in order to be Hamiltonian.
 Whereas to be Lagrangian, the system must also satisfy the assumption of motion equivalence.
@@
 
-Nevertheless, based on the equivalence between the tangent bundle and cotangent bundle formulations of Lagrangian and Hamiltonian mechanics, one might wonder whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics \parencites[1182]{Barrett2}.
+%% Not sure if it is worth raising the following worry. If so, need to fix/clarify the response, which is sketchy.
+%Nevertheless, based on the equivalence between the tangent bundle and cotangent bundle formulations of Lagrangian and Hamiltonian mechanics, one might wonder whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics \parencites[1182]{Barrett2}.
 For, if there were, then this formulation of Lagrangian mechanics might be more general than the cotangent bundle formulation of Hamiltonian mechanics.
@@
 For the action seems to be inherently a globally-defined quantity, as an integral defined over a trajectory.
-
-[To do: figure out best place to slip in the following background information, which is relevant for the preceding objection:] Barrett shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics are categorically equivalent provided that we define categories of models over the tangent and cotangent bundle formulations, respectively \parencites*[1181-82]{Barrett2}.
-However, he also shows that if we consider the more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are not categorically equivalent \parencites*[1182-83]{Barrett2}.
-This parallels the construction of Hamiltonian mechanics in Section~\ref{derivation}.
 
@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-[At least in a footnote: comment on how this characterization of infinitesimal particles relates to standard philosophy of physics interpretive disputes over the status of point particles in classical mechanics.
-See Butterfield Pointillisme article; Butterfield characterizes particles as extensionless.]
-
-Let $\mathcal{C}$ be the state space for the whole system.
-Any point $c \in \mathcal{C}$ represents a particular state of the whole system.
-We define a classical \textit{particle} as the limit of recursive subdivision.
-A classical system then consists of a collection of classical particles.
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{For discussion of the status of extensionless point particles within classical mechanics, see \textcites[]{ButterfieldPointillisme}.}
+%comment on how this characterization of infinitesimal particles relates to standard philosophy of physics interpretive disputes over the status of point particles in classical mechanics.
+Butterfield characterizes particles as extensionless.
+
+Let $\mathcal{C}$ be the state space for the whole system, so that any point $c \in \mathcal{C}$ represents a particular state of the whole system.
+Given infinitesimal reducibility, a classical system consists of a collection of classical particles.
 Let $\mathcal{S}$ be the state space for the particles, so that each point in $\mathcal{S}$ describes a possible state of a classical particle.
-We assume that this state space $\mathcal{S}$ is a manifold.
-Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
+We assume that this state space $\mathcal{S}$ is a manifold.\footnote{Ultimately, we derive an if-then claim: if the state space of classical particles is a manifold, then it must be a symplectic manifold.} Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
 That is, the state of the whole system is a distribution over its infinitesimal parts.
@@
 We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state.
-Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$, yielding a real-valued vector for any given particle state, with $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates,' but we will reserve that terminology for the state variables that define units.} We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
+Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$; this function yields a real-valued vector for any given particle state, possessing $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates,' but we will reserve that terminology for the canonical position coordinates, which are the state variables that define physical units.} We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
 For instance, $q$ might be position, defining units of meters.
@@
 
+% % Ask GC if instances of `state variables' in the next paragraph are correct; i.e. `state variables' rather than `coordinates'
+
 On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function.
 As a function defined over the state variables, the distribution transforms as a density, i.e.
-$\rho_c$ changes as a function of the Jacobian determinant that relates the new coordinates to the old coordinates.
+$\rho_c$ changes as a function of the Jacobian determinant that relates new state variables to old ones.
 As a function defined over the points in phase space, the distribution transforms as a scalar, i.e.
-$\rho_c$ is invariant under all coordinate transformations.
+$\rho_c$ is invariant under all state variable transformations.
 Hence, the Jacobian determinant must in fact equal one.
-In other words, $\rho_c$ is a function only of the particle state $s$ and does not depend on the choice of coordinates.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\xi^a)$ and $\rho_c(\hat{\xi}^b)$ represent the same distribution as expressed over the different coordinates and $s(\xi^a)$ and $s(\hat{\xi}^b)$ represent the same particle state as expressed by the different coordinates.
-Note that $\rho_c(s(\hat{\xi}^b))$ shows how the distribution depends on the coordinates only through the particle state.} We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
+This means that $\rho_c$ is a function only of the particle state $s$ and does not depend on the choice of coordinates.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\xi^a)$ and $\rho_c(\hat{\xi}^b)$ represent the same distribution as expressed over different state variables and $s(\xi^a)$ and $s(\hat{\xi}^b)$ represent the same particle state as expressed by the different variables.
+Note that $\rho_c(s(\hat{\xi}^b))$ shows how the distribution depends on the state variables only through the particle state.} We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
 In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a density and scalar function.
@@
 
-The argument begins with the simplest case: suppose that a single coordinate $q $ suffices to define the unit system for the particle state space.
+We begin with the simplest case: suppose that a single coordinate $q $ suffices to define the unit system for the particle state space.
 Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number.
@@
 Since---by assumption---$q$ fully defines the unit system, $\hat{q}$ must do so as well, making it a function just of the old coordinate: $\hat{q}=\hat{q}(q)$.
-Call the new units $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
+Call the new state variables $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
 As shown above, the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one.
@@
 Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$.
-Suppose you change the first coordinate $q^1$ without changing the others.
-Then we will find a variable $k_1$ that changes as before so that the densities are invariant.
-Now change the second coordinate $q^2$ in the same way while also fixing $k_1$.
-Then we find a corresponding $k_2$.
+If we change the first coordinate $q^1$ without changing the others, then the variable $k_1$ must change as before, so that the density remains invariant.
+Likewise, transforming the second coordinate $q^2$ in the same way (while fixing $k_1$) requires a corresponding $k_2$.
 In the end, we will find that $\mathcal{S}$ is $2n$-dimensional, and the state variables are $\xi^a = \{ q^i, k_i \}$.
@@
 We can now show that $\mathcal{S}$ must have at least the structure of a symplectic manifold.
-Invariant densities can only be assigned to infinitesimal areas.
-Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface, so that we can define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$, for $\Sigma \subset \mathcal{S}$.
+Defining an invariant density requires the existence of infinitesimal areas.
+Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface.
+This lets us define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$, for $\Sigma \subset \mathcal{S}$ (i.e.
+the fraction of the system whose parts are in the region $\Sigma$ of state space $\mathcal{S}$).
 Because degrees of freedom are---at least locally---independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
@@
 Therefore the form $\omega$ itself cannot be degenerate (for otherwise, we would have states that are not counted by it).
+% % Invariant densities can only be assigned to infinitesimal areas. Get a citation for this claim
+
 Next, we show that $\omega$ must be closed, using the fact that degrees of freedom are independent.
@@
 
+The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is a coordinate-invariant distribution over the state space $\mathcal{S}$ of its infinitesimal parts.
+If the state space of the particles is a manifold, then it must be a symplectic manifold: symplectic manifold structure is necessary for the distribution to be coordinate-invariant.
+% Could connect this with the earlier discussion of North's methodology of seeking coordinate-invariants, from section~\ref{fundamentality}.
+
 %since the degrees of freedom are independent, the number of states on a surface does not change if we translate that state across independent degrees of freedom.
@@
 
-The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is a coordinate-invariant distribution over the state space $\mathcal{S}$ of its infinitesimal parts.
-If the state space of the particles is a manifold, then it must be a symplectic manifold: symplectic manifold structure is necessary for the distribution to be coordinate-invariant.
 % Which is the only type of manifold that supports coordinate-invariant distributions
@@
 
-Next, we assume that the classical systems of interest are deterministic and reversible. This amounts to satisfying the following joint condition:
+Next, we focus on those classical systems that are deterministic and reversible. This amounts to satisfying the following joint condition:
 
@@
 \noindent
-\textit{Deterministic and reversible evolution assumption}: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.
+\textit{Deterministic and reversible evolution}: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.
 
 \end{quotation}
-
-[Should we footnote that this condition leaves out some classical systems, e.g. Norton's dome, Earman's space invaders?]
 
@@
 The evolution is therefore a symplectomorphism and corresponds to Hamiltonian evolution.
+
 Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes (i.e.
@@
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).
-Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.\footnote{Weak equivalence corresponds to a general treatment of Lagrangian mechanics that does not presuppose Riemannian geometry.}
+Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.\footnote{Weak equivalence corresponds to a general treatment of Lagrangian mechanics that does not presuppose Riemannian geometry.
+As we discuss in Section~\ref{Curiel}, Equations~\ref{weak_equivalence} correspond to the kinematical constraints central to Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics.}
 
@@
 \end{equation}
-where V is another arbitrary function. We recognize this as the Hamiltonian for massive particles under potential forces.
-
-[To do: Connect this with Curiel equation 6.2 for the Hamiltonian.]
+where V is another arbitrary function.
+We recognize this as the Hamiltonian for massive particles under potential forces.
+Whereas Curiel \parencites*[]{Curiel} argues that this Hamiltonian requires the imposition of \textit{ad hoc} conditions, we take ourselves to have given an eminently natural construction of it.
+We return to this point in Section~\ref{Curiel}, using it to defuse Curiel's argument for the primacy of Lagrangian mechanics.
+
+This completes our central argument for privileging Hamiltonian mechanics over Lagrangian mechanics.
+We have seen that Lagrangian systems are a proper subset of Hamiltonian systems, specifically those Hamiltonian systems that also satisfy the motion equivalence condition.
+In the remainder of this essay, we provide further reasons for privileging Hamiltonian mechanics.
+The next section describes how Hamiltonian mechanics provides the natural setting for the invariant density, describing the matter content of the theory.
+Section~\ref{fundamental} argues that the invariant phase space density is more fundamental than the state variables, since conjugate momentum is a gauge-dependent quantity.
+We then turn in Section~\ref{Lagrangian} to provide a geometric interpretation of the principle of stationary action within Hamiltonian mechanics.
+Our construction makes manifest the gauge redundancy of the Lagrangian, providing a further reason to privilege the Hamiltonian.
+Finally, we respond to Curiel's arguments against the fundamentality of Hamiltonian mechanics.
 
@@
 
-We have now shown that Hamiltonian mechanics requires less structure than Lagrangian mechanics because it requires fewer physical conditions.
+Hamiltonian mechanics is more fundamental than Lagrangian mechanics because Hamiltonian systems require fewer physical conditions.
 While this provides one reason for privileging Hamiltonian mechanics, it is not the only one.
-The special relationship between Hamiltonian mechanics and invariant densities provides a further---no less important---reason for privileging Hamiltonian mechanics.
+The special relationship between Hamiltonian mechanics and invariant densities provides a further reason for privileging Hamiltonian mechanics.
 In this section, we argue that viewing physics through the lens of invariant densities has much to commend it.
@@
 If the scalar value of the density at one point is larger than another, this entails that the density is higher at that point of phase space.
-Physically, this means that there is more physical stuff at this point.
+Physically, this means that there is more matter at this point.
 Likewise, if the scalar value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
-For brevity, will refer to this as the property of \textit{point-wise comparison}.
+For brevity, we'll refer to this as the property of \textit{point-wise comparison}.
 Point-wise comparison requires a coordinate-invariant density.
-Defining an invariant density requires, in turn, that we define state space over a symplectic manifold, as the derivation in Section~\ref{deterministic} shows.
-In other words, to define a density with the property of point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
+As the derivation in Section~\ref{deterministic} shows, to guarantee the existence of an invariant density, we must define state space over a symplectic manifold.
+In other words, to generically have a density with the property of point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
 
 Section~\ref{dependence} shows why point-wise comparison does not hold in general: the density is a coordinate-dependent quantity, and thus it is not generically coordinate-invariant.
-Consequently, within the Lagrangian formalism, point-wise comparison holds only within special coordinate systems.
+Consequently, within the Lagrangian formalism, point-wise comparison holds only when we have made a special choice of coordinates.
 This gives a further reason to privilege Hamiltonian mechanics.
@@
 Since the density is coordinate-dependent, this shows that coordinate-independence is not a necessary condition for fundamental physical significance.
-Our position undercuts metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms [get citations: Field 1980; Chen papers; North].
-Additionally, we explain how the conjugate momentum $p$ is itself a gauge-dependent quantity, bolstering our claim that the density is fundamental, not the state variables.
-Hence, the real physical significance of canonical coordinates is that they allow us to define an invariant phase space density $\rho $.
+Our position undercuts metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms.\footnote{For arguments in favor of representing physical systems and explanations in exclusively intrinsic terms, see \textcites[]{Field,North,Chen}.} Additionally, we explain how the conjugate momentum $p$ is itself a gauge-dependent quantity, bolstering our claim that the density is more fundamental than the state variable pairs.
+We argue that the real physical significance of canonical coordinates is that they allow us to define an invariant phase space density $\rho $.
 
@@
 The density is coordinate dependent because its definition requires us to make a choice of coordinates.
-Worse than that, it is not even coordinate invariant: the density changes with the Jacobian.
+Worse than that, it is not generically coordinate invariant, since the density changes with the Jacobian.
 For example, changing from polar to Cartesian coordinates, we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
@@
 
-\subsection{Coordinate-dependent yet Fundamental}
+%\subsection{Fundamental density and gauge-redundant momentum}
+\subsection{Coordinate-dependent yet fundamental} %; Coordinate-variant yet significant
 \label{fundamental}
@@
 
-Section~\ref{fundamentality} analyzed fundamentality in terms of generality.
+Section~\ref{fundamentality} characterized fundamentality in terms of generality.
 We argued that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in part because it is more general.
@@
 It quantifies the amount of stuff in each phase space state.
-Section~\ref{deterministic} demonstrated what is perhaps a third reason for viewing the density as fundamental: by defining the density, we recover the right mathematical structure, namely a symplectic manifold.
-If we did not define the density, then we would need another explanation of why this mathematical structure arises for classical systems.
+
+Section~\ref{deterministic} demonstrated a third reason for viewing the density as fundamental: by defining the density, we recover the right mathematical structure, namely a symplectic manifold.
+If we did not define the density, then we would need another account for why this mathematical structure arises for classical systems.
+Collectively, these considerations speak in favor of viewing the invariant phase space density as not only physically significant but also fundamental.
 %see 11:50am of 12-21 meeting video for clarification of this point. There are probably finer details for me to incorporate.
 
-Collectively, these considerations speak in favor of viewing the invariant phase space density as not only physically significant but also fundamental.
 Yet, as Section~\ref{dependence} explained, the density is always coordinate-dependent, even when coordinate-invariant.
@@
 Affording the density fundamental significance therefore places us at odds with those who take coordinate-independence as a necessary condition for fundamentality.
-[To do: provide literature references for this and motivating quotations]
-
-Some philosophers have instead privileged the state variables $(q, p)$.
+Instead, we take coordinate-invariance as a weaker necessary (but not sufficient) condition for fundamental physical significance.
+% % [To do: provide literature references for this and motivating quotations; likewise for earlier remark about intrinsicality: Field, North, Chen]
+
+% % To do: insert discussion of stat mech and entropy here: This is why statistical mechanics works best on phase space (we have a uniform measure over q and p).
+Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
+Areas/volumes are measures of number of states.
+They are coordinate invariant.
+Also the Shannon and Gibbs entropies are coordinate invariant.
+
+Being necessary for fundamental significance does not, however, make coordinate-invariance necessary for \textit{non-fundamental} physical significance.
+Some quantities and properties are coordinate-variant yet still physically significant.\footnote{Indeed, many physically significant quantities are gauge-redundant, such as conjugate momentum and the Lagrangian (see Sections~\ref{gauge_momentum} and~\ref{gauge}).
+Insofar as gauge-redundancy is analogous to being coordinate-variant, this supplies a further reason for denying that coordinate-invariance is a necessary condition for physical significance.} Our position here again places us at odds with others in the literature.
+For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
+However, some choices of canonical coordinates $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [to do: provide a clear example of this].
+Similarly, we saw in Section~\ref{dependence} that some coordinate choices in Lagrangian mechanics are physically significant because they enable the property of point-wise comparison.
+This is related to the fact that, in Lagrangian mechanics, Cartesian coordinates are a physically significant coordinate choice because they decompose position and velocity variables as independent degrees of freedom.
+This independence does not hold generically across coordinate systems.\footnote{In contrast, canonical position and conjugate momentum variables always characterize separable/independent degrees of freedom in the Hamiltonian framework.} Finally, as the derivation in Section~\ref{deterministic} shows, a necessary condition for defining the symplectic structure is that we can express the density in a particular choice of units.
+If we did not have the physical ability to define the density in some units we would not be able to define this structure.
+This provides another instance of a physically significant but coordinate-dependent representational choice.
+Indeed, the next section shows that the conjugate momentum is
+
+% % Could insert a summary paragraph either here or at the end of the next subsection explaining the following: coordinate-invariance and gauge-invariance are both necessary for fundamental physical status.
+But they're not necessary for being physically significant.
+Both coordinate-variant and gauge-redundant quantities can still be physically significant, although they cannot be physically fundamental.
+So in this way, we view gauge-redundancy as a kind of coordinate-variants (rather than a kind of coordinate-dependence).
+And gauge-invariance is analogous to a kind of coordinate-invariance, rather than a kind of coordinate-independence.
+And perhaps this helps explain the physical significance of gauge-variant quantities.
+% Question: is any quantity that is coordinate-variant automatically coordinate-dependent?
+If so, then coordinate-variances a special case of coordinate-dependence?
+And so gauge-variance would also be a special case of coordinate-dependence?
+
+\subsection{Gauge redundancy of conjugate momentum}
+\label{gauge_momentum}
+
+%The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density (as the fundamental description of a classical state)?
+
+Whereas we privilege the invariant phase space density, some philosophers have instead privileged the canonical state variables $(q, p)$.
 For instance, in her argument for the fundamentality of Hamiltonian mechanics, North claims that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
-[To do: incorporate Wallace 2003 (p.
-164) claim that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
-See Curiel 2014, ft 16].
-Yet, as we show below, the conjugate momentum $p$ is actually a gauge-redundant quantity.
+Likewise, \textcites[164]{Wallace2003} claims that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
+Yet, as we now show, the conjugate momentum $p$ is actually gauge-redundant.
 Assuming that gauge-invariance is a necessary condition for fundamental physical significance, this precludes $p$ from having fundamental status.
@@
 In this way, the invariant density is more physical than the state variables themselves.
+% % Double check Wallace quotation. See Curiel 2014, ft 16]. 
 % Possibly a stronger claim: This entails that the invariant density is really the important physical element/object in Hamiltonian mechanics.
-
-[To do: explain why we have grounds to privilege the invariant density over not only the conjugate momenta but also the canonical position coordinate.
-Is this connected with the following thought: Hamiltonian mechanics does not really care what the q's and p's are.
-It is the motion assumption that tells us that q's are position, simply because they define the unit system.
-]
-
-[To do: think about whether treating gauge-invariance as a necessary condition for fundamentality is in tension with our position, specifically our contention that coordinate-dependency does NOT preclude a quantity from fundamental physical significance.
-Does our position require an important physical difference between gauge-redundancy and coordinate-dependence?].
-% % Note that if coherent, our position has the following consequence: gauge-redundancy is a sufficient condition for a quantity to be non-fundamental.
-But coordinate-dependency is NOT a sufficient condition for being non-fundamental.
-This could strike some readers as in tension.
-If we are comfortable viewing gauge-redundancy as precluding a quantity from fundamental significance, then why not coordinate-dependence?
-
-We now consider transformations that leave the density, position, and velocity invariant.
+% % Could also connect standard interpretation of the canonical state variable pair with the earlier quotation from Ruetsche, as these being fundamental in the physicist sense.
+Shows how we are providing a more fine-grained physical notion of fundamentality, by going beyond just the epistemic criterion that Ruetsche provides.
+Both the invariant density and the canonical state variable pair are fundamental in the physicist sense.
+But the invariant density is more fundamental because it does not involve gauge-redundancy.
+
+To show that the conjugate momentum is gauge-redundant, consider the set of transformations that leave the density, position, and velocity invariant.
 We will see that we can interpret these transformations as a set of gauge transformations on the conjugate momentum $p $.
-This entails that---unlike both the canonical position state variable $q$ and the kinetic momentum---the conjugate momentum is a gauge-dependent quantity.
+This entails that the conjugate momentum is a gauge-dependent quantity, unlike both the position state variable $q$ and the \textit{kinetic} momentum.
 Recall that motion equivalence entails the following: knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
@@
 The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
-Additionally, it shows that not all state variables are on equal physical footing.
+Insofar as coordinate-invariance is a plausible necessary condition for fundamental physical status, so is gauge-invariance.
+By being gauge-variant, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can remain physically significant but not fundamental).
+
+Additionally, the gauge redundancy of conjugate momentum shows that not all state variables are on equal physical footing.
 Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.
-Indeed, as we now argue, some state variable choices are more physically significant than others.
+To see this, recall that the phase space density remains invariant upon changing the coordinates $q$, inducing a state variable transformation on the conjugate momenta $p$ (which compensate by transforming covariantly).
+It is in this sense that the coordinates $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.\footnote{As discussed in Section~\ref{fundamental}, these units are physically significant despite not being coordinate-invariant.} Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are of equal physical significance.
+In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
+
 % Whereas symplectic geometry creates a mathematical, syntactic symmetry between the $q$ and $p$ variables, this is not a physically significant symmetry from the standpoint of measurement.
@@
 
-In privileging the phase space density, we have relied on its coordinate-invariance.
-Plausibly, coordinate-invariance is a necessary condition for fundamental physical significance.
-Being necessary for fundamental significance does not, however, make coordinate-invariance necessary for physical significance simpliciter.
-Our position here again places us at odds with others in the literature.
-For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
-Yet, some choices of the coordinates $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [to do: provide a clear example of this].
-The phase space density remains invariant upon changing the coordinates $q$, inducing a state variable transformation on the conjugate momenta $p$ (which compensate by transforming covariantly).
-It is in this sense that the coordinates $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.
-These units are physically significant despite not being coordinate-invariant.
-Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are physically significant.
-In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative  $p$'s are not.
-
-[Possible further point to include: discuss how Cartesian coordinates can be physically significant, in virtue of position and velocity being independent degrees of freedom, whereas this independence does not hold generically across coordinate systems.
-(In contrast, position and conjugate momentum variables always characterize separable/independent degrees of freedom).]
-
-[Not sure if/where to include the following point (possibly with the discussion of the coordinate-dependence of the density?): As the derivation in Section~\ref{deterministic} shows, a necessary condition for defining the symplectic structure is that we can express the density in a particular choice of units.
-If we did not have the physical ability to define the density in some units we would not be able to define this structure.]
-
-[To do: expand on the following point:] This is why statistical mechanics works best on phase space (we have a uniform measure over q and p).
-Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
-Areas/volumes are measures of number of states.
-They are coordinate invariant.
-Also the Shannon and Gibbs entropies are coordinate invariant.
+% % [To do: explain why we have grounds to privilege the invariant density over not only the conjugate momenta but also the canonical position coordinate.
+Is this connected with the following thought: Hamiltonian mechanics does not really care what the q's and p's are.
+It is the motion assumption that tells us that q's are position, simply because they define the unit system.
+]
+% Possible response: the invariant density by itself can specify the state of the classical system.
+Whereas the canonical position coordinate (although fundamental in its own right) is insufficient for defining the state.
+So it is only the pair of canonical state variables that could define the state, but we have problematized the fundamental status of the conjugate momentum.
+
+% % [To do: think about whether treating gauge-invariance as a necessary condition for fundamentality is in tension with our position, specifically our contention that coordinate-dependency does NOT preclude a quantity from fundamental physical significance.
+Does our position require an important physical difference between gauge-redundancy and coordinate-dependence?].
+Maybe the thought is that gauge-invariance and coordinate-invariance are on the same footing.
+So if coordinate-invariance is a necessary condition for physical significance, then so is gauge-invariance.
+A gauge-redundant quantity can still be physically significant, in the same way that a coordinate -dependent quantity can be physically significant.
+% % Note that if coherent, our position has the following consequence: gauge-redundancy is a sufficient condition for a quantity to be non-fundamental (in the same way that coordinate-variance is sufficient for non-fundamentality).
+But coordinate-dependency is NOT a sufficient condition for being non-fundamental.
+This could strike some readers as in tension.
+If we are comfortable viewing gauge-redundancy as precluding a quantity from fundamental significance, then why not coordinate-dependence?
+Response: we just need to motivate that gauge-redundancy is more analogous to coordinate-variance than coordinate-dependence.
 

### 2021-02-08 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Rewrote Curiel, conclusion, density
Added in references as well and further edits to every section. Search XXX for major remaining to do items.

--- 2021-02-07 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 
+% % following command suppresses ``in" for articles and inproceedings only
+\renewbibmacro{in:}{%
+\ifboolexpr{%
+test {\ifentrytype{article}}%
+or
+test {\ifentrytype{inproceedings}}%
+}
+{}%<-nothing
+{\printtext{\bibstring{in}\intitlepunct}}%<-normal ’in’ with punctuation
+}
+
 \newcommand{\diff}{\mathop{}\!\mathrm{d}} %defines exterior derivative 
@@
 
-The intellectual richness of classical mechanics manifests itself in an abundance of formulations: with Newtonian, Hamiltonian, Lagrangian, and Hamilton--Jacobi being the most prominent.
-Butterfield \parencites*[]{Butterfield} has aptly referred to these formulations as different \textit{frameworks} for treating classical systems.
+The intellectual richness of classical mechanics manifests itself in an abundance of formulations, with Newtonian, Hamiltonian, Lagrangian, and Hamilton--Jacobi being among the most prominent.
+Butterfield \parencites*[]{Butterfield2006} has aptly referred to these formulations as different \textit{frameworks} for treating classical systems.
 Recent work in philosophy of science has focused largely on how these formulations relate to one another, with the aim of arriving at a deeper understanding of classical mechanics.
-By considering the physical and conceptual differences between these formulations, we can better distinguish physical features from artifacts of our representations.
+By considering the conceptual differences between these formulations, we can better interpret the physical significance of our representations.
 
@@
 Section~\ref{derivation} describes in detail the derivation of Hamiltonian mechanics for systems that satisfy two physical conditions: infinitesimal reducibility and deterministic/reversible evolution.
-We then show (Section~\ref{motion}) that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
+We then show that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
 This completes our primary argument for privileging Hamiltonian mechanics.
 
-Subsequent sections provide additional reasons to privilege Hamiltonian mechanics, along with illustrating further interpretive benefits of our approach.
+Subsequent sections describe additional advantages of Hamiltonian mechanics, along with further interpretive benefits of our approach.
 Section~\ref{density} explains how the invariant phase space density plays a special role in the Hamiltonian framework.
-The density allows us to compare and track the amount of material in different regions of phase space.
+The density allows us to compare and track the amount of matter in different regions of phase space.
 In contrast, Lagrangian mechanics does not support these important properties of the density.
@@
 Each condition entails physically significant properties, described in the middle column.
-Each physical property corresponds to a formal mathematical framework depicted in the right column: including phase space, Hamiltonian mechanics, and Lagrangian mechanics.
+Each physical property corresponds to a formal mathematical framework depicted in the right column, including phase space, Hamiltonian mechanics, and Lagrangian mechanics.
 In this way, our framework develops a correspondence between physical conditions and their associated mathematical formalisms.
@@
 In the widest sense, a classical system is any infinitesimally reducible system.
-Colloquially, we can say that the whole state is equal to the sum of its parts, where we take these infinitesimal parts to be classical particles.
+Colloquially, we can say that the whole state equals the sum of its parts, where we take these infinitesimal parts to be classical particles.
 
@@
 The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
-
-[To do: Clarify the scope of our treatment: at least in this paper, it seems to be restricted to $n$-particle classical mechanics.
-Does it also handle rigid bodies and classical continuum mechanics?
-Cite Mark Wilson 2013 on these issues].
 % % I think I could spend more time clarifying reducibility, infinitesimal. Maybe define separately first?
@@
 Hamiltonian mechanics picks out the subset of Newtonian systems that follow deterministic and reversible evolution.
-Our treatment thereby sets aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \parencites*[]{Earman,Baez,Norton}.} Additionally, we set aside dissipative systems, i.e.
+Our treatment thereby sets aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \textcites*[3-4]{Baez}{Earman}{Norton}.} Additionally, we set aside dissipative systems, i.e.
 those that fail to conserve energy.
@@
 Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
+
+[XXX To do: Clarify the scope of our treatment: our discussion makes it seem restricted to $n$-particle classical mechanics.
+Does it also handle rigid bodies and classical continuum mechanics?
+Cite Mark Wilson \parencite*[]{Wilson} on these issues].
+
 %To avoid verbal disputes, it is essential to clarify what falls within the scope of ``classical systems." One way to delimit classical systems is to specify a set of equations that they satisfy---such as Hamilton's equations or the Euler--Lagrange equations.
 We will instead follow a different approach by delimiting classical systems based on the physical conditions that they satisfy.
-
- %[To do: on the topic of failures of determinism, cite Earman  \textit{a primer on determinism}, Baez singularities paper, or Norton's dome paper].
 
@@
 \textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{fundamental} discusses further senses in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
-Compared with North's \parencites*[]{North} account of structure, our definition has two distinct advantages.
+Compared with North's \parencites*[]{North} account of structure, our definition has two advantages.
 First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which empiricists and pragmatists decry.
@@
 
-Despite our different commitments, our approach has many methodological parallels to North's.
+Despite our different commitments, our approach has methodological parallels to North's.
 We agree with North that it is fruitful to identify both invariants and ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}.
 Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North}.
-Indeed, we undertake precisely this sort of investigation in Section~\ref{derivation}.
-However, North advocates a metaphysical account of structure, about which we remain agnostic.
+Indeed, we undertake precisely this sort of investigation in Section~\ref{derivation}.\footnote{Curiel also aims to identify the necessary and sufficient mathematical structures for formulating classical mechanics.
+He sets out to establish the ``geometrical structure necessary (and manifestly sufficient) for the formulation of the Euler--Lagrange equation'' relevant for all kinematically-possible solutions \parencites*[292]{Curiel}.
+See Section~\ref{Curiel} for discussion.} However, North advocates a metaphysical account of structure, about which we remain agnostic.
 According to North, ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" \parencites*[66]{North}.
@@
 Regardless of its metaphysical status, Hamiltonian mechanics is physically more fundamental, and we will argue for its epistemic and conceptual primacy as well.
-Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [to do: mention these different Hamiltonian formalisms and provide references].
+Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [XXX to do: mention these different Hamiltonian formalisms and provide references].
 Each of these frameworks has different pragmatic costs and benefits.
@@
 As they note, Galilean boosts are often symmetry transformations of Lagrangian systems (in the sense of preserving the equations of motion), but they are not point transformations.
-Additionally, the two-body (Kepler) problem admits canonical transformations that map physically distinct solutions to each other (specifically, solutions whose orbits possess the same energy but different eccentricity).\footnote{For more details on the classical two-body problem, see \textcites[]{Belot}.} Hence, we cannot straightforwardly compare the structure of Lagrangian and Hamiltonian mechanics in terms of the symmetry groups of their statespaces.
+Additionally, the two-body problem admits canonical transformations that map physically distinct solutions to each other (specifically, solutions whose orbits possess the same energy but different eccentricity).\footnote{For more details on the classical two-body problem, see \textcites[]{Belot}.} Hence, we cannot straightforwardly compare the structure of Lagrangian and Hamiltonian mechanics in terms of the symmetry groups of their statespaces.
 
@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{For discussion of the status of extensionless point particles within classical mechanics, see \textcites[]{ButterfieldPointillisme}.}
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{For discussion of the status of extensionless point particles within classical mechanics, see \textcites[]{Butterfieldpoints}.}
 %comment on how this characterization of infinitesimal particles relates to standard philosophy of physics interpretive disputes over the status of point particles in classical mechanics.
@@
 trajectories in physical space-time) suffices to recover its dynamics (i.e.
-evolutions in state space) and vice-versa.\footnote{ \textcites[]{Abraham1978}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time (see \textcites[1180-1181]{Barrett2} for discussion).
+evolutions in state space) and vice-versa.\footnote{ \textcites[]{Abraham}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time (see \textcites[1180-1181]{Barrett2} for discussion).
 Motion equivalence can be understood as the converse of this theorem: assuming that Lagrangian and Hamiltonian mechanics agree on particle evolutions, we derive that the Legendre transformation is well-defined in the hyper-regular domain.}
@@
 If we integrate, we have:
-\begin{equation} 
+\begin{equation} \label{linear_relationship}
 \begin{aligned}
@@
 where V is another arbitrary function.
-We recognize this as the Hamiltonian for massive particles under potential forces.
+We recognize this as the Hamiltonian for massive particles under potential forces, covering both classical electrodynamics and classical Newtonian gravitation.
+For example, $A$ can play the role of the magnetic vector potential, with $V$ playing the role of the electric potential.
 Whereas Curiel \parencites*[]{Curiel} argues that this Hamiltonian requires the imposition of \textit{ad hoc} conditions, we take ourselves to have given an eminently natural construction of it.
@@
 Hamiltonian mechanics is more fundamental than Lagrangian mechanics because Hamiltonian systems require fewer physical conditions.
-While this provides one reason for privileging Hamiltonian mechanics, it is not the only one.
-The special relationship between Hamiltonian mechanics and invariant densities provides a further reason for privileging Hamiltonian mechanics.
-In this section, we argue that viewing physics through the lens of invariant densities has much to commend it.
+While this provides one reason to privilege Hamiltonian mechanics, it is not the only one.
+In this section, we argue that the special relationship between Hamiltonian mechanics and invariant densities provides a further reason to privilege Hamiltonian mechanics.
+Additionally, we argue that viewing physics through the lens of invariant densities has much to commend it.
 In general, it provides a better understanding of classical systems.
@@
 Since the density is coordinate-dependent, this shows that coordinate-independence is not a necessary condition for fundamental physical significance.
-Our position undercuts metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms.\footnote{For arguments in favor of representing physical systems and explanations in exclusively intrinsic terms, see \textcites[]{Field,North,Chen}.} Additionally, we explain how the conjugate momentum $p$ is itself a gauge-dependent quantity, bolstering our claim that the density is more fundamental than the state variable pairs.
-We argue that the real physical significance of canonical coordinates is that they allow us to define an invariant phase space density $\rho $.
+Our position thereby undercuts metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms.\footnote{For arguments in favor of representing physical systems and explanations in exclusively intrinsic terms, see \textcites[]{Field,North,Chen2018}.} Section~\ref{gauge_momentum} shows why the conjugate momentum $p$ is itself a gauge-dependent quantity.
+This bolsters our claim that the density is more fundamental than state variable pairs.
+It also shows that---\textit{pace} conventional wisdom---position and conjugate momentum are \textit{not} on equal physical footing within Hamiltonian mechanics.
+% Canonical coordinates are physically significant because they allow us to define an invariant phase space density.
+% % This latter claim might seem puzzling, since insofar as we need to use canonical coordinates to define the density, this seems to speak in favor of the canonical coordinates being more fundamental.
 
@@
 \label{dependence}
+% densities as coordinate -dependent and variant
 
@@
 More precisely, if $\rho(x, y)$ is a function of multiple variables, dimensional analysis requires that the units of $\rho$ are consistent with the units of the arguments in the expression $\int_U \rho(x, y) dx dy$.
-In polar coordinates $(r, \theta)$, if we measure $r$ in meters and $\theta$ in radians, then $\rho(r, \theta)$ must return mass over meter-radians .
-
-The density is coordinate dependent because its definition requires us to make a choice of coordinates.
+In polar coordinates $(r, \theta)$, if we measure $r$ in meters and $\theta$ in radians, then $\rho(r, \theta)$ must return mass over meter-radians.
+
+Hence, the density is coordinate dependent because its definition requires us to make a choice of coordinates.
 Worse than that, it is not generically coordinate invariant, since the density changes with the Jacobian.
@@
 This makes the value of the density, in general, not directly comparable at different points.
-Hence, to determine whether there is more stuff at one location in state space than another, we need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
+Hence, to determine whether there is more stuff at one location in state space than another, we generically need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
 we can't just compare the values of $\rho$ in these coordinates and see which is larger).
@@
 [see 57.3 minutes of 12-21 recording for discussion]
+
+The Hamiltonian formalism avoids this further epistemic requirement by defining the phase space density in terms of quantities that vary covariantly.
+Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
+In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
+% % Try to precisely gloss/specify the condition for a coordinate system to be special in this way, such that it supports point--wise comparison.
+(Seems inappropriate to mention the Jacobian, since this matters for transforming from one coordinate system to another.)
+
+In response to this deficiency of Lagrangian mechanics, one might wonder if it is possible to construct a coordinate-independent density.
+Section~\ref{independence} considers two strategies for doing so, showing why both fail.
+The first appeals to volume forms from differential geometry.
+The second appeals to particular features of Riemannian geometry, which suffices for the treatment of many Lagrangian systems.
+In each case, we will see that the coordinate-dependence of the physically significant density $\rho$ is unavoidable.
+Due to this coordinate-dependence, the density is in general not coordinate-invariant.
+Readers satisfied with the discussion so far can proceed to Section~\ref{comparing}, which illustrates one interpretive benefit of focusing on the density: it provides a better understanding of the relationship between Hamiltonian and Lagrangian mechanics within the hyper-regular domain.
+
+\subsubsection{A coordinate-independent density?}
+\label{independence}
 
@@
 This is only possible if the determinant of $g_{ab}$ equals one.
-Hence, we must define $\hat{\rho} $ in a special coordinate system, namely one where---at least locally---$|g_{ab}| = 1$.
+Hence, we must define $\hat{\rho} $ in a special coordinate system, namely one where $|g_{ab}| = 1$.
 For this reason, $\hat{\rho} $ remains coordinate-dependent.
+%det g_ab =1 means cells you construct are equal volume. whereas cartesian means not just equal volume but sides (like cubes)
 %Fortunately, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
@@
 
-We can now look at what happens for distributions over position and velocity, i.e.
+\subsubsection{Comparing densities on the tangent vs. cotangent bundles}
+\label{comparing}
+
+% % Wonder if it is possible to move this material to the `interpreting Lagrangian mechanics' section?
+
+Focusing on the density has the further benefit of elucidating the relationship between Hamiltonian and Lagrangian mechanics, especially within the hyper-regular domain.
+Consider a distribution over position and velocity, i.e.
 on the tangent bundle in Lagrangian mechanics.
@@
 \wedge dv^n = |g_{ab}| \bigwedge dx^i \bigwedge dv^j$.
-Hence, the determinant of the metric tensor is equals the component of the volume form of the tangent bundle.
+Hence, the determinant of the metric tensor equals the component of the volume form of the tangent bundle.
 The relationship to the density as expressed in phase space is therefore the following:
@@
 
-To summarize the main theme from this section: in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
-In this context, we need to know which coordinate system we are in before making these comparisons.
-The Hamiltonian formalism circumvents this further epistemic requirement by defining the density using quantities that vary covariantly.
-Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
-
 % Remember: the density is coordinate-invariant but not coordinate-independent; it is actually coordinate-dependent (i.e. it does not admit of something like a coordinate-free definition)
@@
 
-% % To do: insert discussion of stat mech and entropy here: This is why statistical mechanics works best on phase space (we have a uniform measure over q and p).
+[XXX To do: insert discussion of stat mech and entropy here: This is why statistical mechanics works best on phase space (we have a uniform measure over q and p).
 Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
@@
 They are coordinate invariant.
-Also the Shannon and Gibbs entropies are coordinate invariant.
+Also the Shannon and Gibbs entropies are coordinate invariant.]
 
@@
 For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
-However, some choices of canonical coordinates $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [to do: provide a clear example of this].
+However, some choices of canonical coordinates $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [XXX to do: provide a clear example of this].
 Similarly, we saw in Section~\ref{dependence} that some coordinate choices in Lagrangian mechanics are physically significant because they enable the property of point-wise comparison.
@@
 For instance, in her argument for the fundamentality of Hamiltonian mechanics, North claims that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
-Likewise, \textcites[164]{Wallace2003} claims that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
+Likewise, \textcites[164,171]{Wallace} claims that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
 Yet, as we now show, the conjugate momentum $p$ is actually gauge-redundant.
@@
 In this way, the invariant density is more physical than the state variables themselves.
-% % Double check Wallace quotation. See Curiel 2014, ft 16]. 
+% % after Double checking Wallace citation, I should probably find a better/more direct one. e.g. Earman article in that volume? 
 % Possibly a stronger claim: This entails that the invariant density is really the important physical element/object in Hamiltonian mechanics.
@@
 This raises a number of objections to our own position, which we rebut.
-In short, we argue that we have given an eminently natural construction of Lagrangian mechanics, viewed as a special case of Hamiltonian mechanics.
+In short, we  argue that we have given an eminently natural construction of Hamiltonian mechanics.
+%In short, we argue that we have given an eminently natural construction of Lagrangian mechanics, viewed as a special case of Hamiltonian mechanics.
 Moreover, our argument shows that Lagrangian mechanics is in some sense less natural than Hamiltonian mechanics because the former is contingent on having a bijection between velocity and conjugate momentum.
@@
 We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating the generalization, see \textcites[110]{Arnold}.
-[To do: verify that Arnold can legitimately be read as generalizing the argument below.]}
+[XXX To do: verify that Arnold can legitimately be read as generalizing the argument below.]}
 % Note on the generalization: cannot use curl.
@@
 Next, we define the phase space evolution over an affine parameter $s $.
-In this extended Hamiltonian framework, the Hamiltonian constraint functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [to do: spell out what the `Hamiltonian constraint' is in general].
+In this extended Hamiltonian framework, the Hamiltonian constraint functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [XXX to do: spell out what the `Hamiltonian constraint' is in general].
 Crucially, the Hamiltonian constraint is now coordinate-invariant, even with respect to time-dependent coordinate transformations.
@@
 This formalism thereby provides a clear physical meaning for the Hamiltonian constraint.
-[To do: provide some reference for this modified Hamiltonian formulation]
-
-[To do: not sure how best to end this subsection/transition to the next one].
+[XXX To do: provide a reference for this modified Hamiltonian formulation].
 We have now considered many conceptual and physical advantages of the Hamiltonian formalism.
-In Section~\ref{Curiel}, we turn to considering Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.
+In the next section, we rebut Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.
+
+%[To do: not sure how best to end this subsection/transition to the next one].
 
@@
 
-Curiel \parencites*[]{Curiel} argues that Hamiltonian mechanics is less physically significant than Lagrangian mechanics.
-He attempts to show that the Lagrangian framework represents classical mechanical systems in a \textit{natural} manner whereas the Hamiltonian formulation involves an artificial construction \parencites*[270]{Curiel}.\footnote{Curiel defines `natural' in terms of natural coordinate systems, defined using the almost-tangent structure of the tangent bundle \parencites*[290-291]{Curiel}.
-He interprets this as capturing the fundamental kinematical constraints relating position and velocity in classical mechanics.} Central to Curiel's argument is the claim that Lagrangian mechanics naturally encodes the kinematical structure of classical systems whereas Hamiltonian mechanics requires imposing an \textit{ad hoc} condition.
-We will now argue that the condition Curiel claims is ad hoc actually has a natural physical interpretation.
-Additionally, we consider a further disagreement about the proper definition of classical systems, which bears on which formal models are interpreted as physically significant.
-% Additionally, we provide reasons to prefer our definition of classical systems over Curiel's, by considering classical systems outside the hyper-regular domain.
-
-Similar to both us and North, one of Curiel's main goals is to identify the necessary and sufficient mathematical structures for formulating classical mechanics.
-Curiel argues that ``the kinematically-possible evolutions" of a classical system correspond exactly to ``the possible solutions to the Euler--Lagrange equation posed on the tangent bundle" of the system's configuration space \parencites*[292]{Curiel}.
-Hence, he sets out to establish the ``geometrical structure necessary (and manifestly sufficient) for the formulation of the Euler--Lagrange equation having those [kinematically-possible] vector fields as its family of solutions" \parencites*[292]{Curiel}.
-Ultimately, Curiel argues that Lagrangian mechanics respects the necessary and sufficient kinematic constraints that any classical system must satisfy \parencites*[307-308]{Curiel}.
-In contrast, Curiel argues that Hamiltonian mechanics does not inherently respect these kinematical constraints.
-Essentially, the kinematical constraints are given by velocity being the first temporal derivative of position: $v^i = d_t x^i = v^i(q^j, p_k)$ (equation~\ref{weak_equivalence}).\footnote{Curiel formalizes this kinematical constraint using the geometrical notion of an almost-tangent structure \parencites*[290]{Curiel}.} Curiel contends that this constraint is `kinematical' in the sense that it must hold independently of the interactions between a system and its environment.
-He interprets it as a substantive physical condition, rather than an analytic truth \parencites*[282]{Curiel}.
-%---interpreted as applying to those systems who possess a unique solution to the Euler--Lagrange equation---. But, Curiel is comfortable talking about Lagrangian systems that are not hyper-regular.
-
-Instead, to treat classical systems according to Hamiltonian mechanics, Curiel argues that we necessarily have to focus on a proper subset of all possible Hamiltonians.
-This subset treats conjugate momentum as a linear function of velocity: $p_i = \lambda^j_i \dot{q}_j $ \parencites*[304]{Curiel}.
-Curiel seems to take this relationship between momentum and velocity as (physically) necessary for a system to be classical.
-It entails that the Hamiltonian takes the following quadratic form---where the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q_i$:
+Curiel argues that Lagrangian mechanics represents classical mechanical systems in a \textit{natural} manner whereas Hamiltonian mechanics involves an essentially artificial construction \parencites*[270]{Curiel}.\footnote{Curiel characterizes `natural' in terms of natural coordinate systems, defined using the almost-tangent structure of the tangent bundle \parencites*[290-291]{Curiel}.
+He also uses the almost-tangent structure to give a general characterization of what he takes to be the fundamental kinematical constraints.} According to Curiel, all physically-meaningful classical systems satisfy a certain set of kinematic constraints.
+In the simplest case, these constraints characterize velocity as the first temporal derivative of position: $v^i = d_t x^i = v^i(q^j, p_k)$ (equation~\ref{weak_equivalence}).\footnote{Curiel treats this constraint as `kinematical' in the sense that it must hold independently of the interactions between a system and its environment.
+He interprets it as a substantive physical condition, rather than an analytic truth \parencites*[282]{Curiel}.} Curiel argues that Lagrangian mechanics encodes the necessary and sufficient conditions for a system to satisfy these kinematic constraints, both within and outside the hyper-regular domain \parencites*[307--308, 311]{Curiel}.
+In contrast, Hamiltonian mechanics does not automatically encode these constraints.
+Since Curiel views these constraints as an essential feature of physically-meaningful classical systems, he argues that Lagrangian mechanics picks out the proper class of classical systems.
+
+Imposing these kinematic constraints within the Hamiltonian formalism requires focusing on a proper subset of all possible Hamiltonians.
+Within this subset, conjugate momentum is a linear function of velocity: $p_i = \lambda^j_i \dot{q}_j $ \parencites*[304]{Curiel}.
+This relationship between momentum and velocity entails that the Hamiltonian takes the following quadratic form---where the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q_i$:
 \begin{equation} \label{6.2}
@@
 \end{equation}
-Curiel takes this form as the ``necessary and sufficient condition [a Hamiltonian must have] for it to model an abstract classical system" \parencites*[305]{Curiel}.
-Nevertheless, Curiel notes that we can formally apply Hamiltonians that do not have the form of Equation~\ref{6.2}, e.g.
-the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$.
-Curiel then argues that all of the ``formally acceptable Hamiltonians" that do not satisfy the quadratic form (\ref{6.2}) are neither physically significant nor physically meaningful.
-It is in this sense that Curiel believes we must artificially restrict Hamiltonian mechanics to those Hamiltonians that satisfy Equation~\ref{6.2}, in order to model physically-meaningful classical systems.
-
-We can thereby reframe Curiel's argument as follows: all physically-meaningful classical systems satisfy a certain set of kinematic constraints, e.g.
-$v^i = d_t x^i$.
-Lagrangian mechanics encodes the necessary and sufficient conditions for a system to satisfy these kinematic constraints, both within and outside the hyper-regular domain.
-In contrast, Hamiltonian mechanics does not encode these kinematical constraints.
-Instead, it applies to non-regular systems that are not classical systems (in virtue of not satisfying these constraints).
-According to Curiel, these systems are not physically meaningful.
-Hence, in order to make Hamiltonian mechanics physically-meaningful, we must impose an ad hoc condition by hand (Equation~\ref{6.2}), since Hamiltonian mechanics itself does not force us to make this restriction.
-
-In response, we will defend two points.
-First, we have in fact already in Section~\ref{motion} seen a natural interpretation of both the linear relationship between momentum and velocity and the quadratic Hamiltonian form of equation~\ref{6.2}.
-Both follow from imposing the motion equivalence condition.
-In equation 5, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (which is connected to the gauge-redundancy of the conjugate momentum).
-Just as Curiel does, we then used this relationship to derive a quadratic form for the Hamiltonian (Equation 7).
+Curiel contends that this form is the ``necessary and sufficient condition [a Hamiltonian must have] for it to model an abstract classical system" \parencites*[305]{Curiel}.
+Since many possible Hamiltonians do not take this form, Curiel believes we must artificially restrict Hamiltonian mechanics to those that satisfy Equation~\ref{6.2}.\footnote{As an example, Curiel considers the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$  \parencites*[305]{Curiel}.}
+%In contrast, Curiel believes that Lagrangian mechanics does not require an analogous restriction.
+
+To rebut Curiel's argument, it suffices to provide a natural interpretation of (\ref{6.2}).
+We have in fact already provided one in Section~\ref{motion}.
+There, we saw how both the linear relationship between momentum and velocity and the quadratic form (\ref{6.2}) follow from imposing the motion equivalence condition.
+In Equation~\ref{linear_relationship}, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (connected to the gauge-redundancy of the conjugate momentum and the electromagnetic field).
+Just as Curiel does, we then used this relationship to derive a quadratic form for the Hamiltonian (Equation~\ref{massive_Hamiltonian}).
 Imposing motion equivalence thereby allows us to interpret this quadratic form as the Hamiltonian for massive particles under potential forces.
-It is not an ad hoc condition imposed on Hamiltonian systems.
+It is not an \textit{ad hoc} condition imposed on Hamiltonian systems.
 Instead, it arises in all contexts where motion equivalence holds, i.e.
 where the trajectory of the system in physical space-time is necessary and sufficient for its evolution in state space.
-
+Thus, we take ourselves to have responded to Curiel's main criticism of Hamiltonian mechanics, namely that it must impose \textit{ad hoc} the condition (\ref{6.2}).
 % Note that our equation 2 (\ref{weak_equivalence}) is a special case (or less mathematically sophisticated formulation) of Curiel's kinematic conditions, i.e.
@@
 
-Secondly, we believe there are some physically-meaningful non-regular Hamiltonian systems.
-These include models of a classical photon.
+Furthermore, Curiel's definition of `classical systems' leads to interpretive challenges that our framework avoids.
+His prohibition against Hamiltonians that violate (\ref{6.2}) excludes Hamiltonians for non-hyper-regular systems.
+However, we believe that there are physically-meaningful non-regular Hamiltonian systems.
+These include classical models of a massless particle moving at a constant speed, such as a classical photon.
 In general, Hamiltonian systems outside the hyper-regular domain correspond to cases where we cannot express the density of states in terms of position and velocity.
-These are cases where the weak equivalence condition (equation~\ref{weak_equivalence}) fails.
+These are cases where the weak equivalence condition (\ref{weak_equivalence}) fails.
 Nevertheless, we can still define a distribution over the phase space points and consider its evolution.
 We thereby view the broader scope of Hamiltonian mechanics as an advantage, rather than a disadvantage.
-This interpretation is compatible with the claim that many---perhaps even most---non-regular Hamiltonian systems are physically uninteresting.
-%[To do: briefly expand on this and connect to Curiel page 309-310 discussion of a similar system]
+Moreover, our interpretation remains compatible with the claim that many---perhaps even most---non-regular Hamiltonian systems are physically uninteresting.
+%could footnote Curiel pages 309-310 discussion of Hamiltonians for systems with constant velocity. see longer comment note below.
+
+Whereas we can easily interpret non-hyper-regular Hamiltonians, the same cannot be said for all non-hyper-regular Lagrangians.
+For example, consider the case where the Lagrangian $L$ equals zero.
+This Lagrangian gives rise to a Hamiltonian that also equals zero.
+By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.
+For this ``system," all paths have the same action, so all paths are possible.
+This gives an initial probability distribution over the states, but it does not specify how this initial distribution will evolve into a future distribution of states.
+In other words, the Lagrangian alone does not supply a probability distribution on the possible evolutions.
+In contrast, the Hamiltonian still defines a unique solution (i.e.
+a unique evolution in phase space) even when outside the hyper-regular domain (i.e.
+when the Hessian is zero).\footnote{In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
+There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.} Indeed, Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}.
+Curiel argues that this is an advantage of Lagrangian mechanics, but it generates challenges for understanding non-hyper-regular Lagrangians.\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.
+In the case where a system is non-regular only at certain points, we can decompose the system into hyper-regular subsystems, each with its own Lagrangian (or Hamiltonian).}
+%The non-hyper-regular points form a boundary that we disregard.
+
+%but he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage.
+On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}.
+In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy the proper kinematical constraints (e.g.
+$v^i = \dot{x}^i$).
+Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will allegedly still satisfy these kinematical constraints \parencites*[311]{Curiel}.
+Curiel therefore views these non-regular, Lagrangian but non-Hamiltonian systems as physically-meaningful classical systems.
+
+Of course, if one insists on defining a `classical system' as Curiel does, one has grounds for privileging Lagrangian mechanics.
+Nevertheless, our definition of classical systems has substantive advantages over Curiel's.
+Preceding sections have already illustrated many of these advantages.
+Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{density}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a more geometrical understanding of the principle of stationary action (Section~\ref{action}).
+Additionally, our definition accommodates models of classical massless particles by maintaining the uniqueness of solutions, even outside the hyper-regular domain.
+Additionally, our definition of a classical system has the advantage of an elegant relationship with quantum mechanical systems.\footnote{For details of the quantum case, see \textcites{AoPPhy1}.
+[XXX: omit this for review].} By changing the requirement of infinitesimal reducibility (Section~\ref{infinitesimal}) to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.
+Hence, our definition perspicuously distinguishes quantum from classical systems.
+We take these considerations as reason enough for adopting our preferred notion of a classical system.
+%Both kinds of systems are Hamiltonian thanks to deterministic and reversible evolution.
 
@@
 Additionally, insofar as this corresponds to the trivial Lagrangian, does this pose a problem for his own endorsement of Lagrangian mechanics outside the hyper-regular domain?
+% in the model of the classical photon, it clearly has a velocity.
+e.g., you always have velocity in Hamiltonian mech.
+one law tells you how velocity is linked to position and velocity.
+So it is not clear what curiel means by saying that non-regular hamiltonian system would not satisfy this kinematic constraint.
+
+% Another reason to privilege our definition: our state space works well for defining the density.
+Whereas using position/velocity leads to the problems with the density that we saw in the preceding section.
+Curiel's framework works well for systems that have inertial frames, Cartesian coordinates, and are closed rather than open.
+Position/velocity variables work well for systems like these, but not if you break one of these conditions.
+In our framework, we have the disadvantage of the gauge dependence of momentum, but this is a small price to pay.
+
+\begin{comment}
+
+% Material that I cut from the Curiel section:
+
+%[To do: briefly expand on this and connect to Curiel page 309-310 discussion of a similar system] 
+% in the model of the classical photon, it clearly has a velocity.
+e.g., you always have velocity in Hamiltonian mech.
+one law tells you how velocity is linked to position and velocity.
+So it is not clear what curiel means by saying that non-regular hamiltonian system would not satisfy this kinematic constraint.
+
+%On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems).
+In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers the Hamiltonian $H=p$, which corresponds to a system moving with constant velocity (e.g.
+velocity of 1).
+So this could very well be a model of a photon, moving at the speed of light.
+Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating the Euler--Lagrange equation.
+Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310].
+But then does Curiel consider this system to not be a classical system?
+Additionally, insofar as this corresponds to the trivial Lagrangian, does this pose a problem for his own endorsement of Lagrangian mechanics outside the hyper-regular domain?
+
+He interprets this as capturing the fundamental kinematical constraints relating position and velocity in classical mechanics.} Central to Curiel's argument is the claim that Lagrangian mechanics naturally encodes the kinematical structure of classical systems whereas Hamiltonian mechanics requires imposing an \textit{ad hoc} condition.
+We will now argue that the condition Curiel claims is ad hoc actually has a natural physical interpretation.
+Additionally, we consider a further disagreement about the proper definition of classical systems, which bears on which formal models are interpreted as physically significant.
+% Additionally, we provide reasons to prefer our definition of classical systems over Curiel's, by considering classical systems outside the hyper-regular domain.
+
+Similar to both us and North, one of Curiel's main goals is to identify the necessary and sufficient mathematical structures for formulating classical mechanics.
+Curiel argues that ``the kinematically-possible evolutions" of a classical system correspond exactly to ``the possible solutions to the Euler--Lagrange equation posed on the tangent bundle" of the system's configuration space \parencites*[292]{Curiel}.
+Hence, he sets out to establish the ``geometrical structure necessary (and manifestly sufficient) for the formulation of the Euler--Lagrange equation having those [kinematically-possible] vector fields as its family of solutions" \parencites*[292]{Curiel}.
+Ultimately, Curiel argues that Lagrangian mechanics respects the necessary and sufficient kinematic constraints that any classical system must satisfy \parencites*[307-308]{Curiel}.
+In contrast, Curiel argues that Hamiltonian mechanics does not inherently respect these kinematical constraints.
+Essentially,
+
+%---interpreted as applying to those systems who possess a unique solution to the Euler--Lagrange equation---. But, Curiel is comfortable talking about Lagrangian systems that are not hyper-regular.
+
+Instead, to treat classical systems according to Hamiltonian mechanics, Curiel argues that we necessarily have to focus on a proper subset of all possible Hamiltonians.
+Nevertheless, Curiel notes that we can formally apply Hamiltonians that do not have the form of Equation~\ref{6.2}, e.g.
+the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$.
+Curiel then argues that all of the ``formally acceptable Hamiltonians" that do not satisfy the quadratic form (\ref{6.2}) are neither physically significant nor physically meaningful.
+It is in this sense that Curiel believes we must artificially restrict Hamiltonian mechanics to those Hamiltonians that satisfy Equation~\ref{6.2}, in order to model physically-meaningful classical systems.
+
+In response, we will defend two points.
+First, we have in fact already in Section~\ref{motion} seen a natural interpretation of both the linear relationship between momentum and velocity and the quadratic Hamiltonian form of equation~\ref{6.2}.
+Both follow from imposing the motion equivalence condition.
+
+% % Another argument one could develop against Curiel: when you move from Newtonian mechanics to a Lagrangian mechanics you appear to lose the ability to describe certain systems.
+A Lagrangian system is fully specified by L(x, v), whereas the Newtonian forces are specified by acceleration as a function of x and v (yieleding three force functions).
+The map between L(x, v) and these forces is a continuous map in the hyper-regular domain, so that a small change in L results in a small change in forces.
+But, outside the hyper-regular domain, there is no unique map from L to these three Newtonian force functions, because there is no single, unique solution.
+So we can write the acceleration as a function of (x, v).
+Whereas in the Newtonian case, we always have a single solution.
+So in other words, the connection between Curiel's equation 4.3 (kinematical constraints defined in terms of the almost tangent structure) vs.
+3.2/3.1 (kinematical constraints in Newtonian mech) is not clear.
+He appears to lose dissipative forces, despite an alleged desire to have Lagrangian mechanics apply to dissipative systems.
+[discussed on 2-1, around 11:25AM]
 
@@
 His definition takes satisfying certain kinematical constraints as fundamental.
-In contrast, we privilege a system having a unique solution, i.e.
+[XXX: state defined at equal time, velocity is not] In contrast, we privilege a system having a unique solution, i.e.
 being deterministic and reversible.
-Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}, but he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage.
-On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}.
-In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy the proper kinematical constraints (e.g.
-$v^i = \dot(x)^i$).
-Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will allegedly still satisfy these kinematical constraints \parencites*[311]{Curiel}.
-Curiel therefore views these non-regular, Lagrangian but non-Hamiltonian systems as physically-meaningful classical systems.
-
-Yet, it is difficult to understand what a Lagrangian system outside the hyper-regular domain could mean.
-For outside this domain, the Euler--Lagrange equations do not have a single, unique solution.
-As an example, consider the case where the Lagrangian $L$ equals zero.
-This Lagrangian corresponds to a Hamiltonian that also equals zero, entailing that $dH = 0 $.
-By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.
-For this ``system," all paths have the same action, so all paths are possible.
-This gives an initial probability distribution over the states, but it does not specify how this initial distribution will evolve into a future distribution of states.
-In other words, the Lagrangian alone does not supply a probability distribution on the possible evolutions.
-In contrast,  a Hamiltonian defines a unique solution (i.e.
-a unique evolution in phase space) even when outside the hyper-regular domain (i.e.
-when the Hessian is zero).\footnote{In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
-There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.}
-
-By providing a clearer conception of classical systems outside the hyper-regular domain, we believe that our definition of a `classical system' has advantages over Curiel's.
-It provides a uniform treatment of phase space distributions, accommodates models of a classical photon, and maintains uniqueness of solutions.
-For this reason, we believe that the burden is on Curiel to clarify the nature of non-hyper-regular Lagrangian systems.
-Additionally, we take ourselves to have responded to Curiel's main criticism of Hamiltonian mechanics, namely that it must impose \textit{ad hoc} the condition (\ref{6.2}).
-We take these considerations as reason enough for adopting our preferred notion of a classical system.
-Of course, if one insists on defining a `classical system' as Curiel does, one will continue to have grounds to privilege Lagrangian mechanics.
-Although we have given reasons to prefer our definition of a classical system, these considerations may not be decisive.\footnote{As described elsewhere, our definition of a classical system has the advantage of an elegant relationship with quantum mechanical systems.
-By changing the requirement of infinitesimal reducibility (Section~\ref{infinitesimal}) to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.
-Hence, our definition elegantly distinguishes between quantum and classical systems.
-Both kinds of systems are Hamiltonian thanks to deterministic and reversible evolution.} Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections.
+
+% % Another reason to privilege our definition/framework for classical system: states should be defined at equal times, i.e.
+at time slices.
+But velocity is not defined at time slices.
+Velocity is defined on a neighborhood.
+So this is a reason to not view velocity as a state variable, pace Curiel.
+It is a reason to prefer phase space when it comes to specifying the state of the system.
+% It is interesting that conjugate momentum but not velocity is defined at a time slice. Insofar as this is true, shows that they are really not trivial notational variants of each other.
+
+% Incorporated the following into a footnote:
+When it comes to analyzing non-hyper-regular systems, there are two cases.
+In the simplest case, the system simply fails to be hyper-regular at certain points.
+We can thereby decompose the system into hyper-regular subsystems, each with its own Lagrangian (or Hamiltonian).
+The non-hyper-regular points form a boundary that we disregard.
+In the more complicated case, the system fails to be hyper-regular \textit{everywhere}.
+It is difficult to interpret a Lagrangian treatment of such systems, since the Euler--Lagrange equations will not have a single, unique solution.
+
+\end{comment}
+
+%Although we have given reasons to prefer our definition of a classical system, these considerations may not be decisive.
+Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections and summarized in the conclusion.
+
+\section{Conclusion}
+\label{conclusion}
+
+By focusing on three basic physical conditions, we have derived the mathematical structures that underpin Hamiltonian and Lagrangian mechanics.
+These three physical conditions are i) infinitesimal reducibility, (ii) deterministic and reversible evolution, and (iii) motion equivalence.
+Each condition gives rise to corresponding mathematical structure, allowing us to interpret this structure in a physically well-motivated fashion.
+By building up the formalism in this way, we have seen that Lagrangian mechanics requires one more condition to hold than Hamiltonian mechanics, namely motion equivalence.
+
+We have used this additional constraint to argue that Hamiltonian mechanics is more general---and hence more fundamental---than Lagrangian mechanics.
+Our argument for privileging Hamiltonian mechanics avoids both the technical problems facing North's \parencites*[]{North} argument and her more substantive metaphysical commitments.
+Likewise, our construction provides a principled response to Curiel's \parencites*[]{Curiel} criticism that Hamiltonian mechanics requires an \textit{ad hoc} condition.
+This argumentative core already provides compelling reasons for interpreting Hamiltonian mechanics as more fundamental than Lagrangian mechanics, despite the fact that both frameworks adequately treat hyper-regular systems.
+
+% As additional reasons in favor of Hamiltonian mechanics, 
+Besides this, our approach has many additional interpretive payoffs.
+Our formalism emphasizes the fundamental significance of the invariant phase space density.
+Whereas the state variables $(q, p)$ suffer from the gauge-redundancy of the conjugate momentum, the density provides an invariant description within the Hamiltonian formalism.
+This explains why statistical mechanics is best done on phase space, since it provides the natural setting for an invariant density.
+We have also elucidated the role of the symplectic form in counting states for each degree of freedom.
+The importance of the density shows that a quantity can be fundamental yet coordinate-dependent.
+We have also argued that non-fundamental quantities can be physically significant despite being neither coordinate- nor gauge-invariant.
+Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain.
+This treatment makes the gauge redundancy of the Lagrangian manifest.
+Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but classical systems in general.
+
+% Note that the momentum is less constrained by the position than velocity is.
+If you are given two positions at different times, then regardless of the evolutions, you can determine the velocity.
+However, from these two positions, you can be told units for the momentum but not what the momentum actually is.
+To do this, you would need to know the right relationship between  x and p.
+For instance, consider a particle that is slowing down: this could be b/c of friction, or gaining mass, or that you're in a comoving frame that is catching up.
+you can't know this just by looking at position.
+you need some additional external knowledge to distinguish these possible scenarios (e.g.
+that mass of the system is constant).
+[Discussed on 2-1, around 11:30AM].
 
@@
 \end{comment}
-
-\section{Conclusion}
-
-To do!
 

### 2021-02-08 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Comments and modifications from meeting

@@
 Despite being coordinate-invariant, the density itself is coordinate-dependent.
-We therefore argue that some quantities are physically significant despite being coordinate-dependent.
+[XXX: one cannot make density truly coordinate independent] We therefore argue that some quantities are physically significant despite being coordinate-dependent.
 Finally, Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework.
@@
 Each condition entails physically significant properties, described in the middle column.
-Each physical property corresponds to a formal mathematical framework depicted in the right column, including phase space, Hamiltonian mechanics, and Lagrangian mechanics.
+Each physical property corresponds to a formal/mathematical structure depicted in the right column, including phase space, Hamiltonian mechanics, and Lagrangian mechanics.
 In this way, our framework develops a correspondence between physical conditions and their associated mathematical formalisms.
@@
 Furthermore, since particle states are independent of a choice of units, this distribution must be invariant.
-We show in Section~\ref{infinitesimal} that to define this invariant distribution, we require a symplectic manifold, which serves as the phase space of classical mechanics.
+We show in Section~\ref{infinitesimal} that defining this invariant distribution requires a symplectic manifold, which serves as the phase space of classical mechanics.
 The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
@@
 In general, knowing a system's position and velocity does not suffice for knowing its momentum.
-Considering a photon as a classical particle, we cannot determine its momentum from its trajectory through space.
-Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
+Considering a photon as a classical particle, we cannot determine its momentum from its trajectory through space.\footnote{XXX: Since photons go at constant speed, one can only recover the direction of momentum from the trajectory, but not the magnitude.
+Or equivalently, two photons may still follow the same trajectory if they only differ in the magnitude of momentum.} Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
 Different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
@@
 Of course, understanding the relationship between Hamiltonian and Lagrangian mechanics partly depends on how one defines ``classical systems." Different definitions can result in different interpretive relationships.
-On our account, Newtonian systems are the widest domain of classical systems, including any system that is infinitesimally reducible.
+On our account, Newtonian systems [XXX: Nope!] are the widest domain of classical systems, including any system that is infinitesimally reducible.
 Hamiltonian mechanics picks out the subset of Newtonian systems that follow deterministic and reversible evolution.
@@
 
+% https://phys.libretexts.org/Bookshelves/Classical_Mechanics/Book%3A_Variational_Principles_in_Classical_Mechanics_(Cline)/10%3A_Nonconservative_Systems/10.04%3A_Rayleighs_Dissipation_Function
+
 [XXX To do: Clarify the scope of our treatment: our discussion makes it seem restricted to $n$-particle classical mechanics.
@@
 Precisely this scenario occurs in the context of Lagrangian mechanics, which requires the additional condition of motion equivalence to hold.
-Whereas North views Lagrangian mechanics as containing ``excess, superfluous structure" \parencites*[75]{North}, we take this additional condition as unnecessary but still physically significant.
+Whereas North views Lagrangian mechanics as containing ``excess, superfluous structure" \parencites*[75]{North}, we take this additional condition as optional but still physically significant.
 Thus, while it is illuminating to identify the minimum mathematical structures necessary for a formulation, additional structure is not automatically rendered physically insignificant.
@@
 
-First, we will show that a classical phase space arises if and only if a system satisfies the following physical condition:
+First, we will show that a classical phase space arises if a system satisfies the following physical condition:
 
@@
 Let $\mathcal{S}$ be the state space for the particles, so that each point in $\mathcal{S}$ describes a possible state of a classical particle.
-We assume that this state space $\mathcal{S}$ is a manifold.\footnote{Ultimately, we derive an if-then claim: if the state space of classical particles is a manifold, then it must be a symplectic manifold.} Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
+We assume that this state space $\mathcal{S}$ is a manifold.\footnote{Ultimately, we derive an if-then claim: if the state space of classical particles is a manifold, then it must be a symplectic manifold.} [XXX: if time is continuous, trajectories in state space must be continuous, local structure must be compatible] Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
 That is, the state of the whole system is a distribution over its infinitesimal parts.
@@
 
-We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state.
-Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$; this function yields a real-valued vector for any given particle state, possessing $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates,' but we will reserve that terminology for the canonical position coordinates, which are the state variables that define physical units.} We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
+We define \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state.
+Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$; this function yields a real-valued vector for any given particle state, possessing $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates,' but we will reserve that terminology for the canonical position coordinates, which are the state variables that define physical units.
+[XXX: coordinates -> unit variables]} We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
 For instance, $q$ might be position, defining units of meters.
@@
 
-On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function.
-As a function defined over the state variables, the distribution transforms as a density, i.e.
+On physical grounds, [XXX: I'd say it's more mathematical grounds] the distribution $\rho_c$ must transform as both a density and a scalar function.
+As a function defined over the state variables [XXX: within the integral $\int_U \rho_c(\xi^a) d\xi^a$], the distribution transforms as a density, i.e.
 $\rho_c$ changes as a function of the Jacobian determinant that relates new state variables to old ones.
-As a function defined over the points in phase space, the distribution transforms as a scalar, i.e.
+As a function defined over the points in phase space [XXX:$\rho_c : \mathcal{S} \to \mathbb{R} $], the distribution transforms as a scalar, i.e.
 $\rho_c$ is invariant under all state variable transformations.
@@
 Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number.
-We will show that, in fact, there can only be a single $k^i$ variable in this case, conjugate to the single $q$ coordinate.
+[XXX: or less] We will show that, in fact, there can only be a single $k^i$ variable in this case, conjugate to the single $q$ coordinate.
 Assuming that the particle density $\rho_c$ is a scalar, it follows that we can arbitrarily change the unit defined by $q$ without changing the density $\rho_c$.
@@
 If the state space of the particles is a manifold, then it must be a symplectic manifold: symplectic manifold structure is necessary for the distribution to be coordinate-invariant.
+Additionally, any symplectic manifold structure will allow the definition of coordinate--invariant distributions: any scalar function can be in fact integrated using the symplectic form in a coordinate-invariant way.
+Therefore the symplectic structure is exactly the structure needed for their definition.
+[XXX:]
 % Could connect this with the earlier discussion of North's methodology of seeking coordinate-invariants, from section~\ref{fundamentality}.
@@
 Since these trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
-If the equations of motions were such that $v^i=v^i(q^i)$, then motion equivalence would fail as the full trajectory would be determined only by the $q^i$.
-Hence, we must have $v^i=v^i(q^j, p_k)$, so that knowledge of the state in phase space suffices for knowledge of the trajectory.
+If the equations of motions were such that $v^i=v^i(q^i)$, then motion equivalence would fail: since both $x^i$ and $v^i$ are a function of only $q^i$, the relationship would not be invertible because of dimensional considerations.
+[XXX] Hence, we must have $v^i=v^i(q^j, p_k)$, so that knowledge of the state in phase space suffices for knowledge of the trajectory.
 If motion equivalence holds, then this relationship must be invertible, meaning that knowledge of the trajectory suffices for knowledge of the phase space state.
@@
 
-Next, we show that this Lagrangian describes massive particles under scalar potentials.
-If we look at the whole distribution (so that $q$ is no longer held fixed), we have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
+Note that weak equivalence is not enough.
+The classical state, as we defined it, is really the composite, the distribution, and we must be able to express that distribution in terms of the position and velocity.
+[XXX] To have full equivalence, we must be able to express the density as well.
+Therefore we must have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
 \begin{equation}
@@
 \end{equation}
-Note that while the value given by $\rho(q^i, p_j)$ is coordinate independent, the value given by $\rho(x^i, v^j)$ depends on the choice of coordinate through $\left|\frac{\partial v^i}{\partial p_j}\right|$.
-If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only.
+Note that while the value given by $\rho(q^i, p_j)$ is invariant under unit, the value given by $\rho(x^i, v^j)$ depends on the choice of units through $\left|\frac{\partial v^i}{\partial p_j}\right|$.
+If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only, since the position variables, by themselves, define the units.
 Similar considerations will also hold for marginal distributions (i.e.
@@
 This bolsters our claim that the density is more fundamental than state variable pairs.
-It also shows that---\textit{pace} conventional wisdom---position and conjugate momentum are \textit{not} on equal physical footing within Hamiltonian mechanics.
+It also shows that---\textit{pace} conventional wisdom---position and conjugate momentum are \textit{not} on equal physical footing within Hamiltonian mechanics.\footnote{This illustrates how one can be mislead by the syntactical form of the mathematics to draw an interpretive conclusion that is not borne out by the physics.[XXX]}
 % Canonical coordinates are physically significant because they allow us to define an invariant phase space density.
@@
 This makes the value of the density, in general, not directly comparable at different points.
-Hence, to determine whether there is more stuff at one location in state space than another, we generically need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
+[XXX: add remark about size of binning with low statistics?] Hence, to determine whether there is more stuff at one location in state space than another, we generically need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
 we can't just compare the values of $\rho$ in these coordinates and see which is larger).

### 2021-02-10 Articles/2019-Entropy/Source/thermodynamics.tex
Another pass at the thermodynamics article

@@
 
-We now want to return to the link between evolution count, state entropy and state spaces.
-In particular, we want to show how this link exists not only in thermodynamics, but also in classical mechanics, quantum mechanics and statistical mechanics.
-This is important not only to show how process entropy relates to the two main ways to define entropy in statistical mechanics, but to give a sense that all geometrical and probabilistic structure assigned to state spaces are ultimately capturing evolution counts in one way or another.
+We now briefly explore the link between evolution count, state entropy and state spaces in other physical theories, namely classical mechanics, quantum mechanics and statistical mechanics.
+This is important not only to show how process entropy relates to the two main ways to define entropy in statistical mechanics, but to give a sense that all geometrical and probabilistic structures assigned to state spaces are ultimately capturing evolution counts in one way or another.
 This is important for our bigger project, Assumptions of Physics, as we want to create a common framework for all physical theories.
 
-We previously defined mechanical systems as those for which all states have the same entropy.
-While one does not usually discuss the entropy of single classical or quantum states, it is easy to see that that must be the case.
+In statistical mechanics there are currently two widely accepted ways to define the entropy of an ensemble.
+The first, commonly referred to as the fundamental postulate of statistical mechanics, is to take the logarithm of the state count of the region of phase space identified by the constraints (e.g.
+the energy of the system where some quantities are within a particular bound).
+One way we can relate this to the process entropy is the following.
+Suppose that the dynamics is Hamiltonian and suppose the constraints are preserved by the dynamics.
+Since each microstate provides a full description of the system, all microstates correspond to the same number of evolutions.
+Therefore the process entropy of the macrostate is, up to a constant, the logarithm of the count of states that satisfy the constraint.\footnote{This is not the only way to relate the two, but it best corresponds to the typical setup one finds in the literature.}
+
+The second way to define entropy is through the Gibbs and von Neumann entropies, which essentially are the application of the Shannon entropy in the classical and quantum case.
+One way we can relate this to the process entropy is the following.
+Suppose we have a macroscopic equilibrium which can be understood as a dynamical equilibrium of the microscopic description.
+That is, the macroscopic equilibrium is a set of evolutions whose microscopic description can be characterized by a stable probability distribution $\rho(x)$.
+An evolution $\lambda(t)$ for the microstate, then, is an dense sequence of infinitely many microstates whose recurrence matches $\rho(x)$.
+Since the state entropy is the maximal process entropy, we are counting all possible evolutions which means all possible permutations.
+The Shannon entropy can be understood as the count of all possible permutations of an infinite sequence, of all possible evolutions.
+
+These links are very straight forward and are fundamental in the sense that we are not applying different definitions that happen to agree: we are trivially applying the general concept, counting evolutions, in the different cases and finding the correct result.
+There is no mystery as to what is going on in the above two paragraphs, it's just a computation.
+
+Note, however, that both definition relies on a somewhat implicit premise: all classical and quantum states individually have the same state entropy, they are, as we defined them, mechanical systems.
 In quantum mechanics, for example, the von Neumann entropy of every pure state is zero.
@@
 So the idea that mechanical systems are associated with a constant entropy, though seldom explicit, is already embedded in the established theories.
+In fact, all geometrical features of those spaces can be understood in terms of counting evolutions.
 
-The geometrical structure of classical mechanics, the symplectic structure of phase space, can be understood to be exactly the structure we need to count evolutions when the evolution is deterministic and reversible and the degrees of freedom are independent.
+The geometrical structure of classical mechanics (i.e.
+the symplectic structure of phase space) can be understood to be exactly the structure we need to count the evolutions of a deterministic and reversible process (i.e.
+a Hamiltonian process).
 In that case state entropy and process entropy coincide, but not in general.
-In dissipative processes, like a damped harmonic oscillator, this relationship breaks down.
-As the system loses energy, the evolutions gradually concentrate around the attractors.
+For example, in dissipative processes (e.g.
+damped harmonic oscillator), this relationship breaks down.
+As the system loses energy, the evolutions gradually concentrate around attractors.
 In terms of count of states, areas shrink over time, so counting states gives us a decreasing number, which is at odds with the idea of entropy increasing when progressing towards an equilibrium.
-What happens is that as the system evolves, each region will correspond to the same number of evolutions, which will now be spread over fewer states: the process entropy (i.e.
+On the other hand, each region as it evolves will correspond to the same number of evolutions, which will now be spread over fewer states: the process entropy (i.e.
 the logarithm of the evolutions per state) increases as evolutions become more concentrated.
+In the same way that thermodynamic entropy is the process entropy under isoentropic evolution (i.e.
+at equilibrium), the state entropy in classical mechanics is the process entropy under isoentropic evolution (i.e.
+deterministic and reversible evolution).
+Therefore, in the same way thermodynamic entropy does not apply outside of equilibrium, one has to be careful to apply the symplectic structure outside of Hamiltonian evolution.
 
@@
 In other words, the inner product tells us how the evolutions coming in and coming out of these processes are to be ``joined'' with the preceding and following ones.
-Moreover, note that these processes are projections, meaning they give the same result if applied more than once.
+Unitary evolution is exactly when all evolution of one state go to one and only one state.
+Moreover, note that these ``black box'' processes are projections, meaning they give the same result if applied more than once.
 Eigenstates of projections, then, are equilibria of the process the projection represents.
@@
 
-In statistical mechanics, one uses the above structures to define the entropy of an ensemble.
-There are currently two ways to do this.
-The first, commonly referred to as the fundamental postulate of statistical mechanics, is to take the logarithm of the state count of the region of phase space identified by the constraints (e.g.
-the energy of the system is within a particular bound).
-We can relate this to the process entropy in the following way.
-Suppose that the constraints are such that each microstate either is compatible or incompatible with the macrostate.
-Then the count of evolutions of the macrostate is the sum of all the evolutions of its microstates.
-Now suppose that, when the constraints are defined, the system is at equilibrium and independent.
-Then the process entropy is equal to the state entropy.
-Since the system is mechanical, the state entropy is the same for all states.
-This means that, if the conditions above are met, the process entropy is, up to a constant, the logarithm of the state count.\footnote{If the system starts independent and at equilibrium, for example, the entropy of a final equilibrium would correspond to the logarithm of the size of the basin of attraction: each state corresponds to a single evolution that ends in the attractor.}
-
-The second way to define entropy is through the Gibbs and von Neumann entropies, which essentially are the application of the Shannon entropy in the classical and quantum case.
-This can be related to the process entropy in the following way.
-Suppose we have a macroscopic equilibrium which can be understood as a dynamical equilibrium of the microscopic description.
-That is, the macroscopic equilibrium is a set of evolutions whose microscopic description can be characterized by a stable probability distribution $\rho(x)$.
-An evolution $\lambda(t)$ for the microstate, then, is an dense sequence of infinitely many microstates whose recurrence matches $\rho(x)$.
-Since the state entropy is the maximal process entropy, we are counting all possible evolutions which means all possible permutations.
-The Shannon entropy can be understood as the count of all possible permutations of an infinite sequence, of all possible evolutions.
-
-The above insights show that the framework we are setting up is not just a nicer retelling of thermodynamics, but it is something more fundamental that underlies all physical theories.
+The above insights show that the concepts we are using do not just provide a nicer retelling of thermodynamics, but they relate to something more fundamental that underlies all physical theories.
+We believe that this is not a coincidence, but stems from an obvious requirement that we must put on physical theories.
 Most, if not all, experiments are described as a sequence of different steps (e.g.
 one or more preparation, manipulation and measurement) stitched together.
-Making sure that the evolution count matches at every junction is a very simple and necessary bookkeeping requirement, and therefore it makes sense it plays such a major role in defining state spaces.
-
-Therefore it makes sense that countMaking sure that the evolution count  Stitching together those phases into a single process
-
-The idea of using states as junctures when processes change is at the core of scientific experimentation since most experiments are
-
- maostate preparation)can help us better understand the fundamental connection between all these theories.
-We stress that we are not arguing that all other notions of entropy should be abandoned in favor of process entropy: quite the contrary.
-Specific tools are often more useful than general tools precisely because they take advantage of the specific nature of the problem.
-The point is that if we are more aware of the interplay between the general definition and the more specific one we will not use the latter inappropriately.
+Therefore a theory must be able to tell us how to ``stitch together'' the evolutions at every junction, how to keep the evolution count correctly.
+This bookkeeping is a requirement and therefore it makes sense it plays such a major role in defining state spaces.
 
@@
 In this paper we gave a conceptual overview for a new approach to recover thermodynamics based on the simple idea of counting evolutions, the possible ways a system can evolve under a process.
-We saw that the logarithm of evolution count under a deterministic process has all the properties we associate with entropy.
-We have seen that the evolution count is maximized at equilibrium and for independent system and only in those cases it depends only on the state.
-A well defined evolution count is necessary if we want to be able to compose processes, since the incoming evolutions needs to match the outgoing evolutions for all states at the junction.
-Thermodynamics studies exactly the case of composite processes of composite systems under some simplifying conditions.
+We saw that the logarithm of evolution count under a deterministic process has all the properties we associate with entropy: it cannot decrease over time, it is constant in reversible processes and it is maximized at equilibrium.
+The state entropy represents the maximized process entropy at equilibrium and for the composite of independent systems.
+Thermodynamic processes and equilibria describe the case where the equilibrium is achieved at all levels: the whole and any of the parts.
 This gives a simple way to rederive thermodynamics in a straight-forward and conceptually insightful way.
 
-Moreover, we have seen that the state spaces of classical, quantum and statistical mechanics include geometrical and probabilistic structures that can all be understood in terms of evolution count.
+We have also seen that two main ways to define entropy in statistical mechanics can be understood as the process entropy for the ensembles, and that the geometrical properties of classical and quantum mechanics can be understood in terms of evolutions counts.
+Keeping track of evolution counts is fundamental in a physical theory for process composition, as we need to make sure that the evolutions are ``stitched together'' correctly when processes are joined.
 Our hope and long term goal is to develop a formal theory based on these ideas that identifies a minimum set of requirements for the definition of a state space.
-This theory can be further refined with additional assumptions in different ways to recover the different physical theories.
+This basic definition can then be further refined with different additional assumptions, each set recovering the different physical theories.
 We believe this would help us better understand the right questions to ask in within the foundations of physics.

### 2021-02-11 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
incorporated edits from Monday 2-8 meeting to first half of paper.

--- 2021-02-08 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 In contrast, Lagrangian mechanics does not support these important properties of the density.
-Despite being coordinate-invariant, the density itself is coordinate-dependent.
+Despite being coordinate-invariant, one cannot make the density truly coordinate independent.
 We therefore argue that some quantities are physically significant despite being coordinate-dependent.
@@
 In general, knowing a system's position and velocity does not suffice for knowing its momentum.
-Considering a photon as a classical particle, we cannot determine its momentum from its trajectory through space.
-Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
+Considering a photon as a classical particle, we cannot determine its momentum from its trajectory through space.\footnote{Since photons travel at constant speed, we can only recover the direction of momentum from the trajectory, but not the magnitude.
+Two photons with different momenta can follow the same trajectory.} Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
 Different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
@@
 
-Of course, understanding the relationship between Hamiltonian and Lagrangian mechanics partly depends on how one defines ``classical systems." Different definitions can result in different interpretive relationships.
-On our account, Newtonian systems are the widest domain of classical systems, including any system that is infinitesimally reducible.
-Hamiltonian mechanics picks out the subset of Newtonian systems that follow deterministic and reversible evolution.
-Our treatment thereby sets aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \textcites*[3-4]{Baez}{Earman}{Norton}.} Additionally, we set aside dissipative systems, i.e.
+Of course, different definitions of ``Hamiltonian" and ``Lagrangian systems" can result in different interpretive relationships.
+On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems satisfy the motion equivalence assumption.
+Since Lagrangian systems satisfy all three conditions (infinitesimal reducibility, deterministic/reversible evolution, and motion equivalence), they constitute the intersection of Newtonian and Hamiltonian systems.} We thereby set aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \textcites*[3-4]{Baez}{Earman}{Norton}.} Additionally, we set aside dissipative systems, i.e.
 those that fail to conserve energy.
-Possessing a deterministic and reversible evolution is equivalent to being non-dissipative, i.e.
-conserving energy.
-Consequently, our derivation of Lagrangian mechanics in Section~\ref{motion} takes place within the context of conservative systems, so we do not consider systems that satisfy inhomogenous Euler--Lagrange equations.
-Although the Lagrangian formulation applies to dissipative systems \parencites[]{Smith,Curiel}, this is not what we will mean by ``Lagrangian systems."\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime.
-We will view non-dissipative systems as Newtonian, but neither Lagrangian nor Hamiltonian.
-Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
-
-[XXX To do: Clarify the scope of our treatment: our discussion makes it seem restricted to $n$-particle classical mechanics.
-Does it also handle rigid bodies and classical continuum mechanics?
-Cite Mark Wilson \parencite*[]{Wilson} on these issues].
+Possessing a deterministic and reversible evolution is equivalent to being non-dissipative.
+Our derivation of Lagrangian mechanics in Section~\ref{motion} takes place within the context of conservative systems, so we do not consider systems that satisfy inhomogenous Euler--Lagrange equations.
+Although both the Lagrangian and Hamiltonian formulations apply to dissipative systems \parencites[\S 10.4]{Cline}, we exclude these.\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
+For details of the Lagrangian approach to dissipative systems, see \textcites[]{Smith}.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime.
+We will view non-dissipative systems as neither Lagrangian nor Hamiltonian.
+Finally, our treatment is not restricted to $n$-particle mechanics but applies as well to continuum mechanics (including rigid bodies and fluid mechanics).\footnote{For discussion of continuum mechanics in the context of classical mechanics, see \textcites[]{Wilson}.} Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
+% For fluid systems, the deterministic/reversible evolution condition applies to the whole system but not for the individual particles due to interactions between these.}
+% % Why don't interactions between particles also arise for $n$-particle mechanics? e.g. Newtonian gravity?
+
+% % XXX: try to find a different citation for the claim that dissipative forces can be incorporated in both Hamiltonian and Lagrangian mechanics. Cline libre text might not look the best!
+
+% https://phys.libretexts.org/Bookshelves/Classical_Mechanics/Book%3A_Variational_Principles_in_Classical_Mechanics_(Cline)/10%3A_Nonconservative_Systems/10.04%3A_Rayleighs_Dissipation_Function
 
@@
 Regardless of its metaphysical status, Hamiltonian mechanics is physically more fundamental, and we will argue for its epistemic and conceptual primacy as well.
-Indeed, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [XXX to do: mention these different Hamiltonian formalisms and provide references].
-Each of these frameworks has different pragmatic costs and benefits.
+Indeed, our argument does not privilege any one of the various mathematical formalizations of Hamiltonian mechanics.\footnote{These include formulating Hamiltonian mechanics on a $2n+1$-dimensional contact manifold, on an extended phase space, on a Poisson manifold, and the standard formulation on a symplectic manifold that we typically use below.} Each of these frameworks has different pragmatic costs and benefits.
 Hence, it is plausible that there may be no single, uniquely necessary mathematical structure for mechanics, but rather a family of related structures.
@@
 This section justifies the physical and mathematical relationships in Figure~\ref{diagram}.
-Sections~\ref{infinitesimal}-\ref{deterministic} show how to derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution.
+Sections~\ref{infinitesimal}-\ref{deterministic} derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution.
 Section~\ref{motion} shows how systems that satisfy the further condition of motion equivalence can be treated within the formalism of Lagrangian mechanics.
@@
 Let $\mathcal{S}$ be the state space for the particles, so that each point in $\mathcal{S}$ describes a possible state of a classical particle.
-We assume that this state space $\mathcal{S}$ is a manifold.\footnote{Ultimately, we derive an if-then claim: if the state space of classical particles is a manifold, then it must be a symplectic manifold.} Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
+It can be shown that this state space must be a manifold, given the continuous evolution in time of the state space variables.\footnote{For details of this derivation, see \textcites[]{AoPPhy1} [omit this for review; or find alternative source].} Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
 That is, the state of the whole system is a distribution over its infinitesimal parts.
@@
 %[is region U in the state space C or S. I assume in S.].
-
-We define the \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state.
-Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$; this function yields a real-valued vector for any given particle state, possessing $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates,' but we will reserve that terminology for the canonical position coordinates, which are the state variables that define physical units.} We will define a \textit{coordinate} $q \in \xi^a$ as a particular state variable that defines a unit.
-For instance, $q$ might be position, defining units of meters.
-In general, the coordinates are the subset of the state variables that define the unit system, needed to measure properties of the system.
+% % XXX could try to find justification/derivation for the state space being a manifold in the Kevin Kelly book, or could cite the math paper with Greenfield; proposition 6.2.
+Try to include a better gloss/motivation for this claim (basic idea: if time is continuous, trajectories in state space must be continuous, local structure must be compatible).
+Curiel paper might also have a brief justification/motivation, if I recall
+
+We define \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state.
+Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$; this function yields a real-valued vector for any given particle state, possessing $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates', terminology that we adopt when speaking about coordinate transformations, coordinate invariance, and coordinate independence.} We will define a \textit{unit variable} $q \in \xi^a$ as a particular state variable that defines a unit.
+For instance, $q$ might be position, defined in units of meters.
+In general, the unit variables are the subset of the state variables that define the unit system, which is needed to measure properties of the system.
+% [XXX: coordinates -> unit variables]
 
 % % Ask GC if instances of `state variables' in the next paragraph are correct; i.e. `state variables' rather than `coordinates'
-
-On physical grounds, the distribution $\rho_c$ must transform as both a density and a scalar function.
-As a function defined over the state variables, the distribution transforms as a density, i.e.
-$\rho_c$ changes as a function of the Jacobian determinant that relates new state variables to old ones.
-As a function defined over the points in phase space, the distribution transforms as a scalar, i.e.
-$\rho_c$ is invariant under all state variable transformations.
-Hence, the Jacobian determinant must in fact equal one.
-This means that $\rho_c$ is a function only of the particle state $s$ and does not depend on the choice of coordinates.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\xi^a)$ and $\rho_c(\hat{\xi}^b)$ represent the same distribution as expressed over different state variables and $s(\xi^a)$ and $s(\hat{\xi}^b)$ represent the same particle state as expressed by the different variables.
+% On physical grounds, [XXX: I'd say it's more mathematical grounds]
+
+Importantly, the distribution $\rho_c$ must transform as both a density and a scalar function.
+As a density, $\rho_c$ is defined over the state variables $\xi^a$, so that we can take integrals of the form $\int_U \rho_c(\xi^a) d\xi^a$ (in this context, $\rho_c:  \mathbb{R}^n \to \mathbb{R}$).
+The density $\rho_c(\xi^a)$ changes as a function of the Jacobian determinant that relates new state variables to old ones.
+As a function defined over the points in phase space ($\rho_c : \mathcal{S} \to \mathbb{R} $), the distribution transforms as a scalar.
+Hence, $\rho_c$ is invariant under all state variable transformations.
+Consequently, the Jacobian determinant must in fact equal one.
+This means that $\rho_c$ is a function only of the particle state $s$ and is invariant under changes of units.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\xi^a)$ and $\rho_c(\hat{\xi}^b)$ represent the same distribution as expressed over different state variables and $s(\xi^a)$ and $s(\hat{\xi}^b)$ represent the same particle state as expressed by the different variables.
 Note that $\rho_c(s(\hat{\xi}^b))$ shows how the distribution depends on the state variables only through the particle state.} We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
@@
 
-We begin with the simplest case: suppose that a single coordinate $q $ suffices to define the unit system for the particle state space.
-Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ can range over more than one number.
-We will show that, in fact, there can only be a single $k^i$ variable in this case, conjugate to the single $q$ coordinate.
-Assuming that the particle density $\rho_c$ is a scalar, it follows that we can arbitrarily change the unit defined by $q$ without changing the density $\rho_c$.
+We begin with the simplest case: suppose that a single unit variable $q $ suffices to define the unit system for the particle state space.
+Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ could be any non-negative integer.
+We will show that, in fact, there must be exactly one $k^i$ variable, conjugate to the single $q$ variable.
 % We further assume that $q$ fully defines the units for all state variables.
@@
 
-Now, suppose that we change the units by transforming the coordinate $q$ to a new coordinate $\hat{q}$.
-Since---by assumption---$q$ fully defines the unit system, $\hat{q}$ must do so as well, making it a function just of the old coordinate: $\hat{q}=\hat{q}(q)$.
+Since the particle density $\rho_c$ is a scalar, we can arbitrarily change the unit defined by $q$ without changing the density $\rho_c$.
+So suppose we change the units by transforming $q$ to a new unit variable $\hat{q}$.
+Since---by assumption---$q$ fully defines the unit system, $\hat{q}$ must do so as well, making it a function of just the old unit: $\hat{q}=\hat{q}(q)$.
 Call the new state variables $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
@@
 But this would mean the unit change cannot be arbitrary.
-Therefore we must have two or more state variables, so that there must be at least one $k^i$ variable.
+Therefore we must have two or more state variables, requiring at least one $k^i$ variable.
 
@@
 Hence, if there were three or more $\xi^a$ variables, this constraint would be insufficient to recover the transformation of the $k^a$ uniquely.
-Yet in this case, $\hat q$  would not fully define the units for all other state variables, which is contrary to assumption.
+Yet in this case, $\hat q$  would not fully define the units for all other state variables, contrary to assumption.
 Consequently, there must be exactly two state variables: $q$ and a single $k$.
-So in this case of a single $q$ variable, coordinate-independent areas and densities can only be defined on a two-dimensional manifold.
+So in the case of a single $q$ variable, defining a coordinate-independent density requires a two-dimensional manifold.
 
@@
 
-To generalize, we say two coordinates are \textit{independent} if changing the units for one does not change the units for the other.
-Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent coordinates $q^i$.
-If we change the first coordinate $q^1$ without changing the others, then the variable $k_1$ must change as before, so that the density remains invariant.
-Likewise, transforming the second coordinate $q^2$ in the same way (while fixing $k_1$) requires a corresponding $k_2$.
+Extending to the general case, note that two unit-variables are \textit{independent} if changing the units for one does not change the units for the other.
+Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent unit-variables $q^i$.
+If we change the first unit-variable $q^1$ without changing the others, then the variable $k_1$ must change as before, so that the density remains invariant.
+Likewise, transforming the second unit-variable $q^2$ in the same way (while fixing $k_1$) requires a corresponding $k_2$.
 In the end, we will find that $\mathcal{S}$ is $2n$-dimensional, and the state variables are $\xi^a = \{ q^i, k_i \}$.
@@
 The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is a coordinate-invariant distribution over the state space $\mathcal{S}$ of its infinitesimal parts.
-If the state space of the particles is a manifold, then it must be a symplectic manifold: symplectic manifold structure is necessary for the distribution to be coordinate-invariant.
+We have shown that the state space of the particles must be a symplectic manifold.
+Additionally, any symplectic manifold structure supports the definition of coordinate-invariant distributions: any scalar function can be integrated using the symplectic form in a coordinate-invariant way.
+Therefore the symplectic structure is exactly the structure needed to define an invariant density.
+Section~\ref{density} describes further interpretive consequences of this connection between invariant densities and symplectic structure.
+% % [XXX: clarify/clean-up this ending.
+also, emphasize/foreshadow the physical signifiance of the distribution and density].
+also, clarify relationship b/w distribution (ontology) and density (epistemic access to/representation of this ontology)
+%Symplectic manifold structure is exactly the structure needed to define an invariant density. 
 % Could connect this with the earlier discussion of North's methodology of seeking coordinate-invariants, from section~\ref{fundamentality}.
@@
 If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.
-Note that each space variable $x^i$ is a coordinate, i.e.
+Note that each space variable $x^i$ is a unit variable, i.e.
 a state variable that defines a unit.
-In fact, the trajectories can be fully described by these units and only these units.
+In fact, the trajectories can be fully described only by these unit variables.
 For, to specify a trajectory, we need epistemic access to a unit system, where the number of independent units is determined by the dimensionality of physical space.
-It follows that $x^i = q^i$, each $q^i$ is paired with a conjugate $p_i$, and each state $\{q^i, p_i\}$ corresponds to one and only one trajectory in physical space.
+It follows that (i) $x^i = q^i$, (ii) each $q^i$ is paired with a conjugate $p_i$, and (iii) each state $\{q^i, p_i\}$ corresponds to one and only one trajectory in physical space.
 At each point $x^i$, then, there must be infinitely many trajectories, one for each possible combination of $\{p_i\}$.
 Since these trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
-If the equations of motions were such that $v^i=v^i(q^i)$, then motion equivalence would fail as the full trajectory would be determined only by the $q^i$.
+If the equations of motion were such that the velocity were a function of only the unit variables (i.e.
+$v^i=v^i(q^i)$), then motion equivalence would fail.
+For if both $x^i$ and $v^i$ were a function of only $q^i$, this relationship would not be invertible because of dimensional considerations (a surjective function from an $n$-dimensional manifold to a $2n$-dimensional manifold cannot be invertible).
 Hence, we must have $v^i=v^i(q^j, p_k)$, so that knowledge of the state in phase space suffices for knowledge of the trajectory.
@@
 In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^i$.
-Therefore, for fixed q, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
+Therefore, for fixed $q$, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
 Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
 Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere, which we take to be positive.
+Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables.
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).
 Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.\footnote{Weak equivalence corresponds to a general treatment of Lagrangian mechanics that does not presuppose Riemannian geometry.
-As we discuss in Section~\ref{Curiel}, Equations~\ref{weak_equivalence} correspond to the kinematical constraints central to Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics.}
-
-Next, we show that this Lagrangian describes massive particles under scalar potentials.
-If we look at the whole distribution (so that $q$ is no longer held fixed), we have $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
+As we discuss in Section~\ref{Curiel}, Equation~\ref{weak_equivalence} corresponds to the kinematical constraints central to Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics.}
+
+Whereas weak equivalence involves motion equivalence at the level of particle states, \textit{full equivalence} extends this to the entire distribution, i.e.
+the composite state of the system.
+Full equivalence holds when we can express the distribution---represented by the density---in terms of position and velocity variables.
+Formally, this requires that $\rho(q^i, p_j) = |J| \rho(x^i, v^j) = \left|\frac{\partial v^i}{\partial p_j}\right| \rho(x^i, v^j)$ since
 \begin{equation}
@@
 \end{equation}
-Note that while the value given by $\rho(q^i, p_j)$ is coordinate independent, the value given by $\rho(x^i, v^j)$ depends on the choice of coordinate through $\left|\frac{\partial v^i}{\partial p_j}\right|$.
-If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only.
-Similar considerations will also hold for marginal distributions (i.e.
-distributions on a subset of the coordinates), which means all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
+Note that while the value given by $\rho(q^i, p_j)$ is invariant under unit changes, the value given by $\rho(x^i, v^j)$ depends on the choice of units through $\left|\frac{\partial v^i}{\partial p_j}\right|$.
+If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only, since the position variables---by themselves---define the units.
+Similar considerations hold for marginal distributions (i.e.
+distributions on a subset of the unit variables), entailing that all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
 We set:\footnote{In Section~\ref{dependence}, we will see that $g^{ij}$ corresponds to the Riemannian metric, since $\partial v^i/\partial p_j$ it is a function only of position.}
@@
 It also shows that---\textit{pace} conventional wisdom---position and conjugate momentum are \textit{not} on equal physical footing within Hamiltonian mechanics.
+The syntactic symmetry between position and momentum in the mathematical formalism does not entail that these quantities are equally physically significant.
+%\footnote{This illustrates how one can be mislead by the syntactical form of the mathematics to draw an interpretive conclusion that is not borne out by the physics.}
 % Canonical coordinates are physically significant because they allow us to define an invariant phase space density.
@@
 
-Hence, the density is coordinate dependent because its definition requires us to make a choice of coordinates.
+Hence, the density is coordinate dependent because its definition requires us to make a choice of unit-variables.
 Worse than that, it is not generically coordinate invariant, since the density changes with the Jacobian.
@@
 This makes the value of the density, in general, not directly comparable at different points.
-Hence, to determine whether there is more stuff at one location in state space than another, we generically need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
+[XXX: mention how this relates to practical considerations governing the choice of bigger bin-sizes when dealing with lower quality statistics] Hence, to determine whether there is more stuff at one location in state space than another, we generically need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
 we can't just compare the values of $\rho$ in these coordinates and see which is larger).
@@
 \end{equation}
-A change of coordinates will induce a contravariant change in the velocity and a covariant change in the momentum.
+A change of unit-variables will induce a contravariant change in the velocity and a covariant change in the momentum.
 The left side will remain invariant.
@@
 This brings us to a second, physical notion of fundamentality: the phase space density $\rho$ is physically fundamental because it tracks how much of the system is in each possible state.
-It quantifies the amount of stuff in each phase space state.
-
-Section~\ref{deterministic} demonstrated a third reason for viewing the density as fundamental: by defining the density, we recover the right mathematical structure, namely a symplectic manifold.
-If we did not define the density, then we would need another account for why this mathematical structure arises for classical systems.
+It quantifies the amount of stuff in each phase space state, in virtue of representing the distribution.
+
+Section~\ref{infinitesimal} demonstrated a third reason for viewing the density as fundamental: by defining the density, we recover the right mathematical structure, namely a symplectic manifold.
+If we did not define the density, then we would need another account of why this mathematical structure arises for classical systems.
 Collectively, these considerations speak in favor of viewing the invariant phase space density as not only physically significant but also fundamental.
@@
 Affording the density fundamental significance therefore places us at odds with those who take coordinate-independence as a necessary condition for fundamentality.
-Instead, we take coordinate-invariance as a weaker necessary (but not sufficient) condition for fundamental physical significance.
+Proponents of this necessary condition might respond by noting that although the density is not coordinate-independent, the distribution is.
+For the distribution is equivalent to the classical state, which is a thing in the world, i.e.
+an ontic entity.
+We grant that the distribution is ontically fundamental in a sense that does not accrue to the density (which represents the distribution).
+However, we maintain that the density itself is of fundamental significance, both epistemically and physically.
+For it is only through the density that we gain epistemic access to the distribution.
+Although any particular choice of coordinates is arbitrary, we need to make some choice in order to quantitatively measure the distribution through experiments.
+Additionally, defining the density is necessary for deriving symplectic structure, as shown in Section~\ref{infinitesimal}.
+Insofar as our opponent views the symplectic structure of Hamiltonian mechanics as fundamental, they must view the density as equally fundamental---despite its coordinate-dependence.
+Likewise, insofar as our opponent refuses to grant the density fundamental status, they will have to withhold this same status from the symplectic structure.
+Hence, the coordinate-independent notion of fundamentality needs to be supplemented with a coordinate-dependent, epistemic notion of fundamentality.
+This latter notion is no less important, and it shows that it is a mistake to desire a completely ``intrinsic" formulation of a physical theory.
+% % Further claim: the ontic, coordinate-independent notion of fundamentality is insufficient for doing physics.
+We need a corresponding epistemic notion of fundamentality, since we units, state variables, and the density in order to connect the distribution to experiment.
+One can argue that any particular choice of state variables or units or coordinates for the density is arbitrary, but we need some arbitrary choice in order to quantitatively measure the distribution through experiments.
+% % Someone might also worry that the density cannot be more fundamental than the state variables, since we define the density over state variables.
+In response, we can say the following: whereas the density is coordinate-invariant, the state variables are not.
+So the state variables are more arbitrary than the density.
+Note that we do not directly measure states of particles.
+We measure quantities.
+So the only way to measure the distribution is to express it as a density in terms of some quantities, i.e.
+in terms of some state variables.
+So we need the state variables to give epistemic access to the distribution, via the density.
+So the state variables themselves are not important (e.g.
+we can choose the unit basically however we want, and the conjugate momentum has even more freedom, due to gauge redundancy).
+The physically important thing is access to the distribution, which we gain through the density.
+The classical system is made up of pieces and the density lets us say how many pieces we have in various states.
+% A defender of the necessity of coordinate-independence for fundamentality might respond as follows: 
 % % [To do: provide literature references for this and motivating quotations; likewise for earlier remark about intrinsicality: Field, North, Chen]
 
-[XXX To do: insert discussion of stat mech and entropy here: This is why statistical mechanics works best on phase space (we have a uniform measure over q and p).
+[XXX To do: insert discussion of statistical mechanics and entropy here: This is why statistical mechanics works best on phase space (we have a uniform measure over q and p).
 Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
@@
 
+Instead, of coordinate-independence, we take coordinate-invariance as a weaker necessary (but not sufficient) condition for fundamental physical significance.
 Being necessary for fundamental significance does not, however, make coordinate-invariance necessary for \textit{non-fundamental} physical significance.
@@
 For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
-However, some choices of canonical coordinates $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [XXX to do: provide a clear example of this].
+However, some choices of unit-variables $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [XXX to do: provide a clear example of this].
 Similarly, we saw in Section~\ref{dependence} that some coordinate choices in Lagrangian mechanics are physically significant because they enable the property of point-wise comparison.
@@
 Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.
-To see this, recall that the phase space density remains invariant upon changing the coordinates $q$, inducing a state variable transformation on the conjugate momenta $p$ (which compensate by transforming covariantly).
-It is in this sense that the coordinates $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.\footnote{As discussed in Section~\ref{fundamental}, these units are physically significant despite not being coordinate-invariant.} Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are of equal physical significance.
+To see this, recall that the phase space density remains invariant upon changing the unit-variables $q$, inducing a state variable transformation on the conjugate momenta $p$ (which compensate by transforming covariantly).
+It is in this sense that the unit-variables $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.\footnote{As discussed in Section~\ref{fundamental}, these units are physically significant despite not being coordinate-invariant.} Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are of equal physical significance.
 In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
@@
 
-Nonetheless, formulating Hamiltonian mechanics within the extended phase space provides one way to rectify these deficiencies.
+Nonetheless, reformulating Hamiltonian mechanics within the extended phase space provides one way to rectify these deficiencies.
 At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and an affine parameter within the usual Hamiltonian framework.
@@
 Next, we define the phase space evolution over an affine parameter $s $.
-In this extended Hamiltonian framework, the Hamiltonian constraint functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [XXX to do: spell out what the `Hamiltonian constraint' is in general].
+In this extended Hamiltonian framework, the Hamiltonian constraint functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [XXX to do: clarify what the `Hamiltonian constraint' is].
 Crucially, the Hamiltonian constraint is now coordinate-invariant, even with respect to time-dependent coordinate transformations.
@@
 This formalism thereby provides a clear physical meaning for the Hamiltonian constraint.
-[XXX To do: provide a reference for this modified Hamiltonian formulation].
+[XXX To do: provide a reference for extened phase space formulation].
 We have now considered many conceptual and physical advantages of the Hamiltonian formalism.
 In the next section, we rebut Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.
+% Is the Hamiltonian constraint for the extended phase space the same as the discussion in this Wikipedia page: https://en.wikipedia.org/wiki/Hamiltonian_constraint
 

### 2021-02-12 Articles/2019-Entropy/Source/thermodynamics.tex
Another pass with Christine. Version sent to Mario.

@@
 To finish, consider a \textbf{deterministic processes with equilibria}, like the one in figure \ref{fig_with_equilibria}.
-An equilibrium for a process is a state that, once it is reached, it is never left.\footnote{This characterization of equilibrium, though amply used in thermodynamic literature, is simplistic and technically does not quite work.
+An equilibrium for a process is a state that, once it is reached, it is never left.\footnote{This characterization of equilibrium, though amply used in thermodynamic literature, is simplistic and technically does not quite work, e.g.
+is not relativistically invariant.
 However, it will suffice for the purpose of this paper as a full treatment would again distract from the main points.} In a process with equilibria all evolutions will reach, at some point, an equilibrium.
@@
 It only knows about evolutions and how to count them.
-The result is completely general and the justification is very straight forward.
-It is conceptually a perfect candidate for the thermodynamic entropy, though we have one questions left: how does it combine under system composition?
+The result is completely general and the justification is very straightforward.
+It is conceptually a perfect candidate for the thermodynamic entropy, though we have one question left: how does it combine under system composition?
 
@@
 This will not happen, however, if the systems are independent.
-That is, as shown in \ref{fig_independence}, for each possible evolution $\lambda_x(t)$ of the first system and $\lambda_y(t)$ of the second system, there will be a possible evolution $\{\lambda_x(t), \lambda_y(t)\}$ on the composite.
+That is, as shown in figure \ref{fig_independence}, for each possible evolution $\lambda_x(t)$ of the first system and $\lambda_y(t)$ of the second system, there will be a possible evolution $\{\lambda_x(t), \lambda_y(t)\}$ on the composite.
 In other words,
@@
 This link between state entropy and independent equilibria opens a series of questions on the fundamental link between the two which go beyond the scope of this paper.
-However, we want at least give a partial answer to one: why is assigning a unique process entropy important?
+However, we want at least to give a partial answer to one: why is assigning a unique process entropy important?
 One answer is process combination.
@@
 If $\lambda(t)$ is an evolution of the combined process $AB$, then for $t \leq t_0$ it must also be an evolution for $A$ and for $t \geq t_0$ it must also be an evolution for $B$.
-This means that at the junction state $x_{t_0} = \lambda(t_0)$ the count of evolution has to match: $\mu_{AB}(x_{t_0})=\mu_A(x_{t_0})=\mu_B(x_{t_0})$.
+This means that at the junction state $x_{t_0} = \lambda(t_0)$ the count of evolutions has to match: $\mu_{AB}(x_{t_0})=\mu_A(x_{t_0})=\mu_B(x_{t_0})$.
 If we want to mix and match from a set of possible processes, then the process entropy at the junction must be only a function of the state, independent of the process.
-As we'll see in section \ref{sec_mechanics}, these type of consideration are already embedded in the mathematical structure of the state spaces of classical mechanics, quantum mechanics, thermodynamics and statistical mechanics.
+As we'll see in section \ref{sec_mechanics}, these types of consideration are already embedded in the mathematical structure of the state spaces of classical mechanics, quantum mechanics, thermodynamics and statistical mechanics.
 
@@
 \begin{itemize}
-	\item all parts of the systems are in equilibria and are independent from each other
+	\item in the final equilibrium, all parts of the system are independent from each other and are themselves in equilibrium
 	\item there is a non-trivial dependence between the initial state of one part and the final state of another (i.e.
@@
 The first requirement tells us that process entropy is maximized at all levels of the system.
-The second requirements serves to rule out the case where the systems are isolated from each other.
+The second requirement serves to rule out the case where the systems are isolated from each other.
 We say a system is in \textbf{thermodynamic equilibrium} if it is part of the output of a thermodynamic process.
@@
 
-The next step is to characterize to what extent can thermodynamic processes change the energy and the entropy of each system.
+For a thermodynamic system, then, the state entropy can be expressed as $S(U, x^i)$ a function of the extensive quantities.\footnote{The existence of unique equilibria between thermodynamic systems forces $S$ to be strictly monotonic in $U$.} This is simply the equation of state in entropic form, which is the starting point of Gibbs' formulation of thermodynamics.
+%[XXX: CITE]
+We can therefore define the relationships
+\begin{align}
+\beta &= \frac{\partial S}{\partial U} = \frac{1}{k_B T} \;\;\;\;\;\; - \beta X_i = \frac{\partial S}{\partial x^i}
+\end{align}
+where $k_B$ is Boltzmann's constant. We find:
+\begin{align}
+dS &= \frac{\partial S}{\partial U} dU + \frac{\partial S}{\partial x^i} dx^i \\
+&= \beta dU - \beta X_i dx^i \\
+dU &= k_B T dS + X_i dx^i
+\end{align}
+which are the usual thermodynamic relationships, except that, for us, the state entropy $S$ is dimensionless.
+
+The next step is to characterize to what extent thermodynamic processes can change the energy and the entropy of each system.
 Intuitively, there are two ``orthogonal'' directions for a change of energy: constant entropy and maximal entropy change.
@@
 We say a thermodynamic system $R$ is a \textbf{thermal reservoir} if the internal energy $U_R$ is the only state variable.
-In this case, the equation of state is simply $S_R = S_R(U_R)$ and $\Delta S_R = \int \beta_R dU_R$.\footnote{The existence of unique equilibria between reservoirs forces $S_R(U_R)$ to be strictly monotonic.} We call \textbf{heat} $Q = - \Delta U_R$ the energy lost by the reservoir during a thermodynamic process.
+In this case, the equation of state is simply $S_R = S_R(U_R)$ and $\Delta S_R = \int \beta_R dU_R$.
+We call \textbf{heat} $Q = - \Delta U_R$ the energy lost by the reservoir during a thermodynamic process.
 
 We say $M$ is a \textbf{mechanical system} if the state entropy is a constant for all states.
-For a thermodynamic system that is also mechanical, we therefore have  $S_M(U_M, x^i_M) = k_M$ for all its states.
+For a thermodynamic system that is also mechanical, we therefore have  $S_M(U_M, x^i_M) = k_M$ for all its states for some constant $k_M$.
 A mechanical system cannot change entropy and all states can be connected through a deterministic and reversible process.
@@
 \begin{aligned}
-dU_M &= k_B T_M dS + X_{Mi} dx^i_M = X_{Mi} dx^i_M.
+dU_M &= k_B T_M dS_M + X_{Mi} dx^i_M = X_{Mi} dx^i_M.
 \end{aligned}
@@
 The energy of the mechanical system is effectively \emph{stored} in one of the other state variables and can be later retrieved.
-We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U_M$ as the energy gained by the system during a thermodynamic process.\footnote{Note that $\frac{\partial S_M}{\partial U_M} = 0 = \beta = \frac{1}{k_B T_M}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite.
+We call \textbf{work} the energy exchanged by a mechanical system and we note $W=\Delta U_M$ as the energy gained by the system during a thermodynamic process.\footnote{Note that $\frac{\partial S_M}{\partial U_M} = 0 = \beta_M = \frac{1}{k_B T_M}$ which means the thermodynamic ``temperature'' of a mechanical system is infinite.
 While this ``temperature'' has little to do with thermometers, the fact that we can always transform all work into heat but not vice-versa is just a specific case of being able to move energy from higher to lower temperature.}
@@
 
-The third law has many equivalent formulations.
-[CITE] One states that the entropy of any system has a non-negative lower bound which is reached by the lowest energy state.
+The third law has many equivalent formulations. %[CITE]
+One states that the entropy of any system has a non-negative lower bound which is reached by the lowest energy state.
 If the ground state is non-degenerate, such as a perfect crystal, its entropy is equal to zero.
@@
 This means that $\beta_X$ is always positive.
-Conversely, $S_X(U_X)$ must be monotonically decreasing, which means $\beta_X$ is always negative.
+Conversely, $S_Y(U_Y)$ must be monotonically decreasing, which means $\beta_Y$ is always negative.
 If we put the two systems in contact, the energy of $Y$ will keep decreasing and the energy of $X$ will keep increasing and no equilibrium will ever be reached.
@@
 No reference to classical mechanics, quantum mechanics or statistical mechanics is needed: the theory stands on its own.
-This fits the original spirit of thermodynamics, which was to find a set of laws that would be independent fn the ultimate dynamics of the constituents.
+This fits the original spirit of thermodynamics, which was to find a set of laws that would be independent of the ultimate dynamics of the constituents.
 This also explains the success of these ideas in fields outside physics.
@@
 That is, the macroscopic equilibrium is a set of evolutions whose microscopic description can be characterized by a stable probability distribution $\rho(x)$.
-An evolution $\lambda(t)$ for the microstate, then, is an dense sequence of infinitely many microstates whose recurrence matches $\rho(x)$.
+An evolution $\lambda(t)$ for the microstate, then, is a dense sequence of infinitely many microstates whose recurrence matches $\rho(x)$.
 Since the state entropy is the maximal process entropy, we are counting all possible evolutions which means all possible permutations.
@@
 
-These links are very straight forward and are fundamental in the sense that we are not applying different definitions that happen to agree: we are trivially applying the general concept, counting evolutions, in the different cases and finding the correct result.
+These links are very straightforward and are fundamental in the sense that we are not applying different definitions that happen to agree: we are trivially applying the general concept, counting evolutions, in the different cases and finding the correct result.
 There is no mystery as to what is going on in the above two paragraphs, it's just a computation.
 
-Note, however, that both definition relies on a somewhat implicit premise: all classical and quantum states individually have the same state entropy, they are, as we defined them, mechanical systems.
+Note, however, that both derivations rely on a somewhat implicit premise: all classical and quantum states individually have the same state entropy; they are mechanical systems as we defined them.
 In quantum mechanics, for example, the von Neumann entropy of every pure state is zero.
@@
 For example, in dissipative processes (e.g.
-damped harmonic oscillator), this relationship breaks down.
+a damped harmonic oscillator), this relationship breaks down.
 As the system loses energy, the evolutions gradually concentrate around attractors.
@@
 the logarithm of the evolutions per state) increases as evolutions become more concentrated.
-In the same way that thermodynamic entropy is the process entropy under isoentropic evolution (i.e.
-at equilibrium), the state entropy in classical mechanics is the process entropy under isoentropic evolution (i.e.
+In the same way that thermodynamic entropy is the process entropy under isentropic evolution (i.e.
+at equilibrium), the state entropy in classical mechanics is the process entropy under isentropic evolution (i.e.
 deterministic and reversible evolution).
-Therefore, in the same way thermodynamic entropy does not apply outside of equilibrium, one has to be careful to apply the symplectic structure outside of Hamiltonian evolution.
+Therefore, in the same way thermodynamic entropy does not apply outside of equilibrium, one has to be careful when trying to apply the symplectic structure outside of Hamiltonian evolution.
 
-The mathematical structure of quantum mechanics also can be understood in terms of evolution counts.
+The mathematical structure of quantum mechanics can also be understood in terms of evolution counts.
 The inner product of quantum mechanics gives us the conditional probabilities of a ``black box'' process (i.e.
@@
 As we saw, conditional probabilities are ratios of evolution counts: what fractions of the evolutions that start in one state end in another.
-In other words, the inner product tells us how the evolutions coming in and coming out of these processes are to be ``joined'' with the preceding and following ones.
-Unitary evolution is exactly when all evolution of one state go to one and only one state.
+In other words, the inner product tells us how the evolutions coming in and going out of these processes are to be ``joined'' with the preceding and following ones.
+Unitary evolution is exactly when all evolutions of one state go to one and only one state.
 Moreover, note that these ``black box'' processes are projections, meaning they give the same result if applied more than once.
@@
 Most, if not all, experiments are described as a sequence of different steps (e.g.
-one or more preparation, manipulation and measurement) stitched together.
+one or more preparations, manipulations and measurements) stitched together.
 Therefore a theory must be able to tell us how to ``stitch together'' the evolutions at every junction, how to keep the evolution count correctly.
@@
 Thermodynamic processes and equilibria describe the case where the equilibrium is achieved at all levels: the whole and any of the parts.
-This gives a simple way to rederive thermodynamics in a straight-forward and conceptually insightful way.
+This gives a simple way to rederive thermodynamics in a straightforward and conceptually insightful way.
 
-We have also seen that two main ways to define entropy in statistical mechanics can be understood as the process entropy for the ensembles, and that the geometrical properties of classical and quantum mechanics can be understood in terms of evolutions counts.
+We have also seen that two main ways to define entropy in statistical mechanics can be understood as the process entropy for the ensembles, and that the geometrical properties of classical and quantum mechanics can be understood in terms of evolution counts.
 Keeping track of evolution counts is fundamental in a physical theory for process composition, as we need to make sure that the evolutions are ``stitched together'' correctly when processes are joined.
@@
 This basic definition can then be further refined with different additional assumptions, each set recovering the different physical theories.
-We believe this would help us better understand the right questions to ask in within the foundations of physics.
+We believe this would help us better understand the right questions to ask within the foundations of physics.
 

### 2021-02-15 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Update classical-phil-article.tex
very minor changes; pushing just so I can pull new changes!

@@
 In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
-If the scalar value of the density at one point is larger than another, this entails that the density is higher at that point of phase space.
+If the value of the density at one point is larger than another, this entails that the density is higher at that point of phase space.
 Physically, this means that there is more matter at this point.
-Likewise, if the scalar value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
+Likewise, if the value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
 For brevity, we'll refer to this as the property of \textit{point-wise comparison}.
 Point-wise comparison requires a coordinate-invariant density.
-As the derivation in Section~\ref{deterministic} shows, to guarantee the existence of an invariant density, we must define state space over a symplectic manifold.
+As the derivation in Section~\ref{infinitesimal} shows, to guarantee the existence of an invariant density, we must define state space over a symplectic manifold.
 In other words, to generically have a density with the property of point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
@@
 Our position thereby undercuts metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms.\footnote{For arguments in favor of representing physical systems and explanations in exclusively intrinsic terms, see \textcites[]{Field,North,Chen2018}.} Section~\ref{gauge_momentum} shows why the conjugate momentum $p$ is itself a gauge-dependent quantity.
-This bolsters our claim that the density is more fundamental than state variable pairs.
+This bolsters our claim that the density is more fundamental than the choice of state variable pairs.
 It also shows that---\textit{pace} conventional wisdom---position and conjugate momentum are \textit{not} on equal physical footing within Hamiltonian mechanics.
@@
 Any and all (time-independent) coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
-See~\ref{gauge} for discussion.} Not only is this formal property methodologically advantageous, but it also bespeaks a further physical significance.
+See Section~\ref{gauge} for discussion.} Not only is this formal property methodologically advantageous, but it also bespeaks a further physical significance.
 This brings us to a second, physical notion of fundamentality: the phase space density $\rho$ is physically fundamental because it tracks how much of the system is in each possible state.
@@
 If we did not define the density, then we would need another account of why this mathematical structure arises for classical systems.
-Collectively, these considerations speak in favor of viewing the invariant phase space density as not only physically significant but also fundamental.
+Collectively, these considerations support interpreting the invariant phase space density as not only physically significant but also fundamental.
 %see 11:50am of 12-21 meeting video for clarification of this point. There are probably finer details for me to incorporate.
@@
 Affording the density fundamental significance therefore places us at odds with those who take coordinate-independence as a necessary condition for fundamentality.
+
 Proponents of this necessary condition might respond by noting that although the density is not coordinate-independent, the distribution is.
@@
 For it is only through the density that we gain epistemic access to the distribution.
-Although any particular choice of coordinates is arbitrary, we need to make some choice in order to quantitatively measure the distribution through experiments.
+Although the choice of coordinates is fairly arbitrary, we need to make some choice in order to quantitatively measure the distribution through experiments.
 Additionally, defining the density is necessary for deriving symplectic structure, as shown in Section~\ref{infinitesimal}.
 Insofar as our opponent views the symplectic structure of Hamiltonian mechanics as fundamental, they must view the density as equally fundamental---despite its coordinate-dependence.
-Likewise, insofar as our opponent refuses to grant the density fundamental status, they will have to withhold this same status from the symplectic structure.
+For the same reason, insofar as our opponent refuses to grant the density fundamental status, they will have to withhold this same status from the symplectic structure.
 Hence, the coordinate-independent notion of fundamentality needs to be supplemented with a coordinate-dependent, epistemic notion of fundamentality.

### 2021-02-18 Articles/2020-QuantumLogic/Source/quantumlogic.tex
Skeleton for quantum logic article

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-1}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{MnSymbol}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\begin{document}

\title{Notes on classical/quantum logic }
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
Notes of classical logic in quantum mechanics.
\end{abstract}

\maketitle

\section{Quantum logic: definition and motivation}

\subsection{Disjunction}

Connection between disjunction and superposition.

``A relevant feature of $\vee$ is that, differently from the case in classical semantics, a quantum disjunction may be true even if neither of it members is true. This reflects, for example, the case in which we are dealing with a state such as that of a spin 1/2 system which is in a linear combination of states up and down. Both propositions, ``the state is up'' and ``the state is down,'' may have no definite truth value (the excluded middle principle is violated), but the disjunction ``the state is up or the state is down'' is a tautology.''

\begin{description}
	\item $p =$ ``The particle has spin $z^+$''
	\item $q =$ ``The particle has spin $z^-$''
\end{description}

Claim: $p \vee q$ is \textbf{always} true, even if $p$ and $q$ may possibly be both false. The idea is that $p \vee q$ includes also all superpositions of $p$ and $q$.

Counterclaim. The propositions $p$ and $q$ are experimentally equivalent to:
\begin{description}
	\item $p =$ ``The expectation value for $S_z$ is 1/2$\hbar$''
	\item $q =$ ``The expectation value for $S_z$ is -1/2$\hbar$''
\end{description}
The proposition $p \vee q$ corresponds to the statement ``The absolute value of the expectation value for $S_z$ 1/2$\hbar$''. If we have a superposition of $p$ and $q$, the expectation value for $S_z$ will be strictly between $-1/2 \hbar$ and $1/2 \hbar$, and therefore $p \vee q$, when discussing expectation values, will be false as one would expect.

\subsection{Distributivity}

The distributivity in quantum mechanics is ``weird'' because (1) it can be true even if none of the elements of the disjunction is true due to the linearity of the Schroendinger equation and (2) quantum observables do not commute.

\begin{description}
    \item $p =$ ``The particle has spin $x^+$''
    \item $q =$ ``The particle has spin $y^+$''
    \item $r =$ ``The particle has spin $y^-$''
\end{description}

Classic: $p \wedge (q \vee r)$ if and only if $(p \wedge  q) \vee (p \wedge r)$

Quantum: Assume $p$ is true, which means we measured spin $x$ to be $+$. Since if I measured spin $y$ I would either get $+$ or $-$, we have $q \vee r \equiv \top$. Therefore $q \vee r \equiv \top$ is true. On the other hand, both terms in the second disjunctions are false: because $p \wedge  q$ refer to incompatible observables, we have $p \wedge  q \equiv \bot$. Therefore the $(p \wedge  q) \vee (p \wedge r) \equiv \bot$.

$q \vee r \equiv \top$ therefore $p \wedge (q \vee r) \equiv p \wedge \top \equiv p$. If $p$ is true

\subsubsection{Reframing}

\begin{description}
    \item $p =$ ``The state of particle at time $t$ is $x^+$''
    \item $q =$ ``The state of particle at time $t$ is $y^+$''
    \item $r =$ ``The state of particle at time $t$ is $y^-$''
\end{description}

Then all three assertions are incompatible with each other. That is, $p \wedge q \equiv q \wedge r \equiv r \wedge p \equiv \bot$. Therefore $p \wedge (q \vee r) \equiv \bot \equiv (p \wedge  q) \vee (p \wedge r)$.

2 casi: considerare

\begin{description}
    \item $p =$ ``The state of particle at time $t$ is $x^+$''
    \item $q =$ ``The state of particle at time $t$ is $y^+$''
    \item $r =$ ``The state of particle at time $t$ is $y^-$''
    \item $\hat{q} =$ ``The state of particle at time $\hat{t}$ is $y^+$''
    \item $\hat{r} =$ ``The state of particle at time $\hat{t}$ is $y^-$''
\end{description}

In this case $p \wedge \hat{q} \nequiv \bot$ and $p \wedge \hat{r} \nequiv \bot$. Therefore the original claim was $p \wedge (\hat{q} \vee \hat{r})$ if and only if $(p \wedge  q) \vee (p \wedge r)$ which is not the classical logic distributive law. Note that, $\hat{q} \vee \hat{r} \equiv \top$ under condition that ``We measure spin $y$ at time $\hat{t}$''.

\subsubsection{Classical example}
\includegraphics[width=\columnwidth]{Balldrop.png}

We have a ball sitting at position $p$ over a hatch. If the hatch opens, the ball drops, bounces around, until it will rest at either position $q$ or $r$ with equal chance. Two cases, as before:

\begin{description}
    \item $p =$ ``The position of the ball at time $t$ is $p$''
    \item $q =$ ``The position of the ball at time $t$ is $q$''
    \item $r =$ ``The position of the ball at time $t$ is $r$''
\end{description}

These are three incompatible assertions. Distributivity holds as before.

\begin{description}
    \item $p =$ ``The position of the ball at time $t$ is $p$''
    \item $q =$ ``The position of the ball at time $t$ is $q$''
    \item $r =$ ``The position of the ball at time $t$ is $r$''
    \item $\hat{q} =$ ``The position of the ball at time $\hat{t}$ is $q$''
    \item $\hat{r} =$ ``The position of the ball at time $\hat{t}$ is $r$''
\end{description}

We assume that, after time $t$, the hatch is opened and that at time $\hat{t}$ the ball has settled. As before, $\hat{q} \vee \hat{r} \equiv \top$ because the ball must be in one of the two positions after it dropped, therefore, if $p$ is true, the left side of the biconditional is true while the right side is false.

\section{Classical and quantum parallel}

Classical state: $\rho(x,p)$

Quantum state: $\psi(x)$

Classical observable $f(x,p)$. Average value $<f> = \int f(x, p) \rho(x, p) dx dp$. Eigenstate $f(x, p) \rho(x, p) = f_0 \rho(x, p)$.

Quantum observable $F$. Average value $<F> = \int \psi^\dagger(x) F \psi(x) dx$. Eigenstate $F \psi(x) = f_0 \psi(x)$.

In both cases, average value equal to $f_0$ means that the contributions of all parts of the ensemble combine to the value $f_0$. Eigenstate $f_0$ means all parts of the ensemble contribute the same value $f_0$.

For example, a classical eigenstate of $H = \frac{p^2}{2m} + \frac{1}{2} k x^2$ with eigenvalue $e_0$ is any function $\rho(x,p)$ that is different from zero only on the ellipses with energy $e_0$. That is, all elements of the distribution have the same energy.

The statement ``the observable $f$ of the system is between $f_0$ and $f_1$'' is therefore unclear. It could be taken to mean ``the average of observable $f$ of the system is between $f_0$ and $f_1$'' or ``the observable $f$ for each part of the ensemble is between $f_0$ and $f_1$''. In the first case, we are identifying a value from the real line, using the standard topology and the standard Borel algebra. In the second case, we are identifying a set of possible functions based on their support.

Now suppose that the $U$ and $V$ are two intervals. We have ``the average $<f>$ is in $U \cup V$'' is  equivalent to ``the average $<f>$ is in $U$'' $\vee$ ``the average $<f>$ is in $V$''. The average value must be in and and only one set. We also have ``the support of $\rho$ is $U \cup V$'' is not equivalent to ``the support of $\rho$ is $U$'' $\vee$ ``the support of $\rho$ is $V$''. The distribution could span both sets at the same time. But the lattice generated by such sets is still a $\sigma$-algebra, and therefore follows classical logic.

To check/prove: neither algebras is a sub-algebra of the other; both sub-algebras of the norm-induced one.

\section{Math}

* We can always use the Borel algebra for our statements. This will include(?) all propositions of ``quantum logic''.

It includes (?) also proposition on expectation values, say of position and momentum. Those will behave exactly like ``point particles''. It also includes (?) propositions on the support.

Show that this is similar to classical distributions.

Claim: in quantum mechanics the ``correct'' logical structure the is algebra of subspaces of the Hilbert space. In this algebra (which is a $\cap$-structure) the meet functions differently than in Boolean logic. Therefore it does not follow classical logic.

Resolution: that algebra does not include many statements that are interesting experimentally. For example, it does not include statements about the expectation values which, in the end, are the result of experiments. (Cross-section?) One

Technical view: the state space of quantum states is a separable Hilbert space; it is a Banach space (complete normed space) and admits a norm-induced topology; it is separable, so it admits a countable basis, so the topology is second countable; therefore the state space has a countably generated Borel-algebra. This, like any set theoretic structure, follows Boolean algebra. Each state is a point and state-statements like ``the system is in state $s_1$'' are singletons in the $\sigma$-algebra. State-statements are all incompatible with each other: either we prepared one state or we prepared another.

If a process with an initial and final state (the internal are disregarded), we have two lattices of propositions. The sample space is the Cartesian product, the $\sigma$-algebra is the product. A measurement process will follow this pattern. We can say ``the system was prepared in state $s$'' and ``the system was measure in state $s$'' and there are two different statements. In the product algebra, we can then mix and match statements with the usual Boolean constructs. Both these are analogous in a classical and quantum case. These will naturally result in different lattices, but they will still be Boolean lattices.

What changes are the probability spaces we can construct. In the quantum case, the probability is not in general a probability measure and has to be consistent with the Born rule (i.e. the inner product).

\bibliography{quantumlogic}

\end{document}

### 2021-02-22 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Light edits and some reference checking
continued attending to existing XXX's.

@@
 We will view non-dissipative systems as neither Lagrangian nor Hamiltonian.
-Finally, our treatment is not restricted to $n$-particle mechanics but applies as well to continuum mechanics (including rigid bodies and fluid mechanics).\footnote{For discussion of continuum mechanics in the context of classical mechanics, see \textcites[]{Wilson}.} Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
+Finally, our treatment is not restricted to $n$-particle mechanics but applies as well to rigid-body and continuum mechanics.\footnote{For discussion of the numerous conceptual problems that arise within different approaches to classical mechanics (including point-masses, rigid-bodies, and continua) see \textcites{Wilson}.
+In Wilson's terminology, the framework we apply here may be yet another `theory facade' for classical physics, but a fruitful one nonetheless.} Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
 % For fluid systems, the deterministic/reversible evolution condition applies to the whole system but not for the individual particles due to interactions between these.}
@@
 
-Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as a formal device for encoding physical information or as a calculational tool.} Whether Hamiltonian phase space structure is metaphysically fundamental (in a classical world) does not matter for the questions we seek to answer.
+Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as a formal device for encoding physical information or as a calculational tool.
+As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is `fundamental in the bottom-layer sense' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental (in a classical world) does not matter for the questions we seek to answer.
 Regardless of its metaphysical status, Hamiltonian mechanics is physically more fundamental, and we will argue for its epistemic and conceptual primacy as well.
@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{For discussion of the status of extensionless point particles within classical mechanics, see \textcites[]{Butterfieldpoints}.}
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{Similarly, Butterfield \parencites[]{Butterfieldpoints} discusses point-particles as being extensionless and Wilson \parencites[]{Wilson} discusses them as being zero-dimensional.}
 %comment on how this characterization of infinitesimal particles relates to standard philosophy of physics interpretive disputes over the status of point particles in classical mechanics.
 Butterfield characterizes particles as extensionless.
+%For discussion of the status of extensionless point particles within classical mechanics, see [have yet to find decent reference for this kind of discussion].
 
@@
 Given infinitesimal reducibility, a classical system consists of a collection of classical particles.
-Let $\mathcal{S}$ be the state space for the particles, so that each point in $\mathcal{S}$ describes a possible state of a classical particle.
-It can be shown that this state space must be a manifold, given the continuous evolution in time of the state space variables.\footnote{For details of this derivation, see \textcites[]{AoPPhy1} [omit this for review; or find alternative source].} Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
+Let $\mathcal{S}$ be the state space for the particles, so that each point $s \in \mathcal{S}$ describes a possible state of a classical particle.
+Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
 That is, the state of the whole system is a distribution over its infinitesimal parts.
-This distribution describes the proportion of classical particles in any particular state.
-Formally, if $\mathcal{S}$ is a manifold, $\int_U \rho_c d\mathcal{S}$ gives us the fraction of the system whose parts are in the region $U$ of the state space $\mathcal{S}$.
+This distribution describes the proportion of classical particles in any particular state from $\mathcal{S}$.
+
+If $\mathcal{S}$ is a discrete space with $U \subseteq \mathcal{S}$, we define $\mu(U) = \sum\limits_{s \in U} \rho_c(s)$ so that $\mu(\mathcal{S})$ gives us the size of the system and $\mu(U)/\mu(\mathcal{S})$ gives us the fraction of the particles that can be found in the region $U$.\footnote{The discrete case is also far less interesting because discrete quantities cannot change continuously over time, meaning that they must be constant.
+For details see \textcites[]{AoPPhy1}.
+[omit this for review]}
 %[is region U in the state space C or S. I assume in S.].
@@
 
-We define \textit{state variables} as a set of quantities $\xi^a$ that fully identify a state.
-Formally, $\xi^a: \mathcal{S} \to \mathbb{R}^n$; this function yields a real-valued vector for any given particle state, possessing $n$-many degrees of freedom.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates', terminology that we adopt when speaking about coordinate transformations, coordinate invariance, and coordinate independence.} We will define a \textit{unit variable} $q \in \xi^a$ as a particular state variable that defines a unit.
+In the continuous case, when $\mathcal{S}$ is a manifold, we define $\mu(U) = \int_U \rho_c(s) d\mathcal{S} $.
+We will call a set of quantities $\xi^a$ \textit{state variables} provided that they fully identify a state, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates'.
+However, we will have multiple manifolds and, to avoid confusion, we will call `state variables' the coordinates of the state space.} We define a density $\rho^{\ast}_c$: $\mathbb{R}^n \to \mathbb{R}$ as the pullback of the distribution $\rho_c$ by the inverse of a coordinate map:  $\rho^{\ast}_c \equiv (\rho_c \circ \xi^{-1})$.
+%To represent the distribution over state variables, we define coordinate charts $(U, \xi)$ over $\mathcal{S}$.
+Likewise, the distribution $\rho_c$ is the pullback of $\rho^{\ast}_c$ by a coordinate map: $\rho_c (s) = (\xi^{\ast} \rho^{\ast}_c)(s) = \rho^{\ast}_c (\xi (s))$.
+
+The distribution $\rho_c$ is a function of the particle states $s \in \mathcal{S}$, while the density $\rho^{\ast}_c$ is a function of the state variables $\xi^a \in \mathbb{R}^n$.
+A coordinate map $\xi$ lets us change variables from $s$ to $\xi^a$.
+Consequently, we require that $\mu(U) = \int_U \rho_c(s) d\mathcal{S} = \int_{\xi(U)} \rho^{\ast}_c (\xi^a) d \xi^a$.\footnote{Indeed, using the Lebesgue measure $\lambda$ on $\mathbb{R}^n$---and assuming that $\xi^{-1}$: $\mathbb{R}^n \to \mathcal{S}$ is measurable---we can define the measure $d\mathcal{S}$ by the pushforward measure $\xi^{-1}_{\ast}(\lambda)(U) \equiv \lambda(\xi(U))$.} Since $\rho_c$: $\mathcal{S} \to \mathbb{R}$ is a function of the particle state $s$, its value is independent of the choice of state variables $\xi^a$.
+Hence, $\rho_c$ is a scalar function and consequently coordinate independent.
+Thus, $\int_U \rho_c(s) d\mathcal{S}$ is invariant under coordinate changes $\xi$ to $\hat{\xi}$.
+As a function of the state variables $\xi^a$, $\rho^{\ast}_c$ transforms according to the Jacobian for this change of variables: $\rho^{\ast}_c (\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho^{\ast}_c (\hat{\xi}^b)$.
+Since $\mu(U) = \int_{\xi(U)} \rho^{\ast}_c (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho^{\ast}_c (\hat{\xi}^b) d \hat{\xi}^b$ remains invariant under this coordinate change, the Jacobian determinant must equal one.
+By dimensional homogeneity, it must also be dimensionless.\footnote{Formally, we can express this argument as a chain of equalities: $\rho^{\ast}_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho^{\ast}_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho^{\ast}_c(\hat\xi^b)$, where $\rho^{\ast}_c(\xi^a)$ and $\rho^{\ast}_c(\hat{\xi}^b)$ are the same density expressed over different state variables and $s(\xi^a) \equiv \xi^{-1}(\xi^a)$ and $s(\hat{\xi}^b)\equiv \hat{\xi}^{-1}(\hat{\xi}^b)$ represent the same particle state as expressed by these different variables.
+Note that $\rho_c(s(\hat{\xi}^b))$ shows how the distribution depends on the state variables only through their labeling of a particle state $s$.}
+%$\rho^{\ast}_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho^{\ast}_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho^{\ast}_c(\hat\xi^b)$
+
+We will now show that in order for the distribution $\rho_c$ to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
+In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a scalar function and for its pullback $\rho^{\ast}_c := \rho_c \circ \xi^{-1}$ to transform as a density.
+
+%In other words, while we can change the units used by the state variables $\xi^a$ to identify the particle states, the units for $\rho^{\ast}_c$ must remain the same so that the density does not change its numeric value
+
+We define a \textit{unit variable} $q \in \xi^a$ as a particular state variable that defines a unit.
 For instance, $q$ might be position, defined in units of meters.
-In general, the unit variables are the subset of the state variables that define the unit system, which is needed to measure properties of the system.
+Two unit variables are \textit{independent} if one unit variable can be changed without changing the other.
+For example, one can measure the vertical position in meters while measuring the horizontal position in miles.
+In general, the unit variables $q^i$ are the subset of the state variables that define the unit system, which is needed to relate the properties of the system to the measured values.
+The idea is that the $q^i$ not only determine the units for themselves, but for all other state variables.
+A \textit{change of units} is therefore a map $\hat{q}^j = \hat{q}^j(q^i)$ that defines new unit variables in terms of the old unit variables alone.
+Since the unit variables define the units for all state variables, this change must induce a unique change of state variables $\hat{\xi}^b = \hat{\xi}^b(\xi^a)$ for all variables.
 % [XXX: coordinates -> unit variables]
@@
 
-Importantly, the distribution $\rho_c$ must transform as both a density and a scalar function.
-As a density, $\rho_c$ is defined over the state variables $\xi^a$, so that we can take integrals of the form $\int_U \rho_c(\xi^a) d\xi^a$ (in this context, $\rho_c:  \mathbb{R}^n \to \mathbb{R}$).
-The density $\rho_c(\xi^a)$ changes as a function of the Jacobian determinant that relates new state variables to old ones.
-As a function defined over the points in phase space ($\rho_c : \mathcal{S} \to \mathbb{R} $), the distribution transforms as a scalar.
-Hence, $\rho_c$ is invariant under all state variable transformations.
-Consequently, the Jacobian determinant must in fact equal one.
-This means that $\rho_c$ is a function only of the particle state $s$ and is invariant under changes of units.\footnote{Formally, we can express this argument as a chain of equalities: $\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$, where $\rho_c(\xi^a)$ and $\rho_c(\hat{\xi}^b)$ represent the same distribution as expressed over different state variables and $s(\xi^a)$ and $s(\hat{\xi}^b)$ represent the same particle state as expressed by the different variables.
-Note that $\rho_c(s(\hat{\xi}^b))$ shows how the distribution depends on the state variables only through the particle state.} We will now show that in order for the distribution to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
-In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a density and scalar function.
 %This is because the distribution is a function only of the states, rather than of the particular choice of coordinates. 
@@
 
-We begin with the simplest case: suppose that a single unit variable $q $ suffices to define the unit system for the particle state space.
-Then, the state variables are given by a set $\xi^a = \{ q, k^i \}$, where \textit{prima facie} the index $i$ could be any non-negative integer.
+Let us consider the simplest case: suppose that a single unit variable $q $ suffices to define the unit system for the particle state space.
+Then, the state variables are given by a set $\xi^a = \{ q, k^\alpha \}$, where \textit{prima facie} the index $\alpha$ could be any non-negative integer.
 We will show that, in fact, there must be exactly one $k^i$ variable, conjugate to the single $q$ variable.
@@
 
-Since the particle density $\rho_c$ is a scalar, we can arbitrarily change the unit defined by $q$ without changing the density $\rho_c$.
-So suppose we change the units by transforming $q$ to a new unit variable $\hat{q}$.
-Since---by assumption---$q$ fully defines the unit system, $\hat{q}$ must do so as well, making it a function of just the old unit: $\hat{q}=\hat{q}(q)$.
-Call the new state variables $\hat{\xi}^b = \{ \hat{q}, \hat{k}^i\}$.
+As a scalar, the units of the distribution $\rho_c$ are invariant under changing the units of the state variables.
+So suppose we change the units by transforming $q$ to a new unit variable $\hat{q}=\hat{q}(q)$.
+Call the new state variables $\hat{\xi}^b = \{ \hat{q}, \hat{k}^\beta\}$.
 As shown above, the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one.
@@
 But this would mean the unit change cannot be arbitrary.
-Therefore we must have two or more state variables, requiring at least one $k^i$ variable.
+Therefore we must have two or more state variables, requiring at least one $k^\alpha$ variable.
+%[XXJ: if previously we measured the density in particles per meter cube---and we change units from meters to miles---won't the units of the density change to particles per cubic mile?
+Isn't this required by dimensional analysis on the expression $\int_U \rho^{\ast}_c(\xi^a) d\xi^a = \int_U \rho^{\ast}_c(\xi^b) d\xi^b$ ?].
 
@@
 Next, we show that in fact there must be exactly one $k$ variable, conjugate to the single $q$ variable.
-Using the fact that the Jacobian determinant equals one, we determine how the state variables $k^i$ must change in order to compensate the change from $q$ to $\hat q$.
-Since $\hat{q}$ depends only on $q$, $\frac{\partial \hat{q}}{\partial k^a} = 0$.
-Hence, the Jacobian is an upper triangular block matrix:
+Using the fact that the Jacobian determinant equals one, we determine how the state variables $k^\alpha$ must change in order to compensate the change from $q$ to $\hat q$.
+Since $\hat{q}$ depends only on $q$, $\frac{\partial \hat{q}}{\partial k^\alpha} = 0$.
+Hence, the Jacobian is a lower triangular block matrix:
 
@@
 |J| = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \begin{vmatrix}
-\frac{\partial \hat{q}}{\partial q} & \frac{\partial \hat{q}}{\partial k^a} \\
-\frac{\partial \hat{k}^b}{\partial q} & \frac{\partial \hat{k}^b}{\partial  k^a} 
+\frac{\partial \hat{q}}{\partial q} & \frac{\partial \hat{q}}{\partial k^\alpha} \\
+\frac{\partial \hat{k}^\beta}{\partial q} & \frac{\partial \hat{k}^\beta}{\partial  k^\alpha} 
 \end{vmatrix} = \begin{vmatrix}
 \frac{\partial \hat{q}}{\partial q} & 0 \\
-\frac{\partial \hat{k}^b}{\partial q} & \frac{\partial \hat{k}^b}{\partial  k^a}
+\frac{\partial \hat{k}^\beta}{\partial q} & \frac{\partial \hat{k}^\beta}{\partial  k^\alpha}
 \end{vmatrix}
@@
 \noindent
-The determinant of a triangular block matrix equals the product of determinants of the block matrices along the diagonal: $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| =  \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right|$.
+The determinant of a triangular block matrix equals the product of determinants of the block matrices along the diagonal: $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| =  \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^\beta}{\partial k^\alpha}\right|$.
 For the Jacobian determinant to equal one, we must therefore have  $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| =   \left|\frac{\partial \hat{q}}{\partial q}\right|^{-1}$.
 Since $\hat q$ is a function only of $q$, $\frac{\partial \hat{q}}{\partial q}$ is a total derivative, so we obtain the inverse by swapping the numerator and denominator.
-Hence, we see that $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$.
+Hence, we see that $\left|\frac{\partial \hat{k}^\beta}{\partial k^\alpha} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$.
 %  $\left|\frac{\partial \hat{q}}{\partial q}\right|^{-1} = \left|\frac{\partial q}{\partial \hat q} \right|$.
 
-This equation provides only a single constraint, placed on the determinant of the transformation from $k^a$ to $\hat{k}^b$.
-Hence, if there were three or more $\xi^a$ variables, this constraint would be insufficient to recover the transformation of the $k^a$ uniquely.
+This equation provides only a single constraint, placed on the determinant of the transformation from $k^\alpha$ to $\hat{k}^\beta$.
+Hence, if there were three or more $\xi^a$ variables, this constraint would be insufficient to recover the transformation of the $k^\alpha$ uniquely.
 Yet in this case, $\hat q$  would not fully define the units for all other state variables, contrary to assumption.
@@
 
-Extending to the general case, note that two unit-variables are \textit{independent} if changing the units for one does not change the units for the other.
+Extending to the general case, recall that two unit-variables are independent if changing the units for one does not change the units for the other.
 Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent unit-variables $q^i$.
@@
 But such a region of state space would---by definition---not be a volume.
-Therefore the form $\omega$ itself cannot be degenerate (for otherwise, we would have states that are not counted by it).
+Therefore the form $\omega$ itself cannot be degenerate (otherwise, some states would not be counted by it).
 % % Invariant densities can only be assigned to infinitesimal areas. Get a citation for this claim
@@
 Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold.
-By convention, we set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
+
+The units we choose for $\omega$ are really the units we choose to count states over an independent degree of freedom.
+This agrees with the practice in statistical mechanics.
+The density $\rho_c$ [XXX: distribution or density?], when normalized, is given in terms of the inverse of the units for $\omega$.
+In accordance with the standard conventions in quantum mechanics and statistical mechanics, the $k_i$ are given in terms of inverse units of the respective $q^i$, such that $dq^i \wedge dk_i$ is a pure number.
+We then use $\hbar$ to define the units for state counting and therefore set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
 % % the `motivation' section of the wiki page on symplectic manifolds provides a different style of derivation for the 2-form being non-degenerate and closed.
+
+In terms of physics, then, a set of canonical coordinates is a set of state variables that express the density in the correct units.
+A canonical transformation is one that preserves the unit and the numerical value of the density.
+A change of unit is a special type of canonical transformation: one where we have an arbitrary change in the unit variables and an induced (covariant) change in the conjugate variables.
 
@@
 trajectories in physical space-time) suffices to recover its dynamics (i.e.
-evolutions in state space) and vice-versa.\footnote{ \textcites[]{Abraham}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time (see \textcites[1180-1181]{Barrett2} for discussion).
+evolutions in state space) and vice-versa.\footnote{ \textcites[218]{Abraham}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time (see \textcites[1180-1181]{Barrett2} for discussion).
 Motion equivalence can be understood as the converse of this theorem: assuming that Lagrangian and Hamiltonian mechanics agree on particle evolutions, we derive that the Legendre transformation is well-defined in the hyper-regular domain.}
@@
 
+%<<<<<<< Updated upstream: XXJ: wasn't sure what this referred to? 
 In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^i$.
 Therefore, for fixed $q$, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
-Since $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
+Let $H$ denote the Hamiltonian.
+Since by Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
 Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere, which we take to be positive.
@@
 In the remainder of this essay, we provide further reasons for privileging Hamiltonian mechanics.
-The next section describes how Hamiltonian mechanics provides the natural setting for the invariant density, describing the matter content of the theory.
+The next section describes how Hamiltonian mechanics provides the natural setting for the invariant density, describing the matter content of the theory [XXX: matter content of the theory seems like an odd phrase].
 Section~\ref{fundamental} argues that the invariant phase space density is more fundamental than the state variables, since conjugate momentum is a gauge-dependent quantity.
@@
 %\footnote{This illustrates how one can be mislead by the syntactical form of the mathematics to draw an interpretive conclusion that is not borne out by the physics.}
+
 % Canonical coordinates are physically significant because they allow us to define an invariant phase space density.
@@
 This relationship shows that the units of $\rho$ depend on the units of both $m$ and $V$.
-More precisely, if $\rho(x, y)$ is a function of multiple variables, dimensional analysis requires that the units of $\rho$ are consistent with the units of the arguments in the expression $\int_U \rho(x, y) dx dy$.
-In polar coordinates $(r, \theta)$, if we measure $r$ in meters and $\theta$ in radians, then $\rho(r, \theta)$ must return mass over meter-radians.
-
-Hence, the density is coordinate dependent because its definition requires us to make a choice of unit-variables.
+The issue is that a manifold, in general, does not provide us with a way to measure areas and volumes.
+If $\rho(x, y)$ is a function of multiple variables, dimensional analysis requires that the units of $\rho$ are consistent with the units of the arguments in the expression $\int_U \rho(x, y) dx dy$.
+For example, a manifold may capture the physical characteristics of a group of people in terms of their height $h$ and their weight $w$, leading to a density $\rho(h,w)$ with units of people/(m$\cdot$Kg).
+Or we may represent these characteristics using height $h$ and body mass index $bmi = w / h^2$, which leads to a different density $\rho(h,bmi)$ with units of (people$\cdot$m)/Kg.
+
+% In polar coordinates $(r, \theta)$, if we measure $r$ in meters and $\theta$ in radians, then $\rho(r, \theta)$ must return mass over meter-radians.
+[changed example.
+Use height/weight to height/body mass index.
+Issue with geometrical examples is that one automatically thinks of geometrical volume.]
+
+Hence, the density is coordinate-dependent because its definition requires us to make a choice of unit-variables.
 Worse than that, it is not generically coordinate invariant, since the density changes with the Jacobian.
-For example, changing from polar to Cartesian coordinates, we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
+In the above example, $\rho(h,w) = |J| \rho(h, bmi) = \frac{1}{h^2} \rho(h, bmi)$.
+For a more familiar example, changing from polar to Cartesian coordinates, we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
 Since the Jacobian is a function of the point, the value of $|J|$ can change from point to point.
@@
 If we have a uniform, constant density in $x$ and $y$, we will have an $r$-dependence in $r$ and $\theta$.
-This dependence is introduced precisely because the area identified by $dr$ and $d\theta$ is $dr \,r d\theta$, which is a function of $r$.
+This dependence is introduced precisely because the region identified by $dr$ and $d\theta$ is $dr \,r d\theta$, which is a function of $r$.
 This makes the value of the density, in general, not directly comparable at different points.
-[XXX: mention how this relates to practical considerations governing the choice of bigger bin-sizes when dealing with lower quality statistics] Hence, to determine whether there is more stuff at one location in state space than another, we generically need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
+Without additional structure, there is no objective sense in which the density is higher at one point or another: it depends on the coordinate.
+
+%there is more stuff at one location in state space than another, we generically need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
 we can't just compare the values of $\rho$ in these coordinates and see which is larger).
@@
 
-The Hamiltonian formalism avoids this further epistemic requirement by defining the phase space density in terms of quantities that vary covariantly.
-Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
+The Hamiltonian formalism fixes this issue by assigning areas to regions in a way that is independent of the choice of unit variables.
+To do so, it forces us to restrict ourselves to canonical coordinates, which explains why momentum has to change covariantly.
+The result is that we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in (in the sense of space-time coordinates).
 In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
@@
 The first appeals to volume forms from differential geometry.
-The second appeals to particular features of Riemannian geometry, which suffices for the treatment of many Lagrangian systems.
+The second appeals to the canonical volume form of Riemannian geometry, which suffices for the treatment of many Lagrangian systems.
 In each case, we will see that the coordinate-dependence of the physically significant density $\rho$ is unavoidable.
@@
 
-In differential geometry, a different mathematical object is called the `density,' namely a volume form (i.e.
+In differential topology, a different mathematical object is called the `density,' namely a volume form (i.e.
 an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation.\footnote{The differential density generalizes volume forms in the case of a non-orientable manifold.
 However, we focus only on cases where the manifold is orientable.
-To simplify the exposition, we will elide subtle differences between volume forms and differential densities.} Our physically significant density $\rho$ is actually the component of this volume form $\nu = \rho(x^i)  \epsilon dx^1 \wedge dx^2 \wedge ...
+To simplify the exposition, we elide subtle differences between volume forms and differential densities, concentrating on the former.} Our physically significant density $\rho$ is actually the component of a volume form $\nu = \rho(x^i)  \epsilon dx^1 \wedge dx^2 \wedge ...
 \wedge dx^n$ where $\epsilon$ is $1$ or $-1$ depending on the orientation.
@@
 the infinitesimal mass contribution we are integrating.
-But this means that the value returned by $\nu$ depends not only on how much stuff there is at a given point, but also on the size of the volume identified by the $n$ vectors given to the form.
+But this means that the value returned by $\nu$ depends not only on how much stuff there is at a given point, but also on the size of the region identified by the $n$ vectors given to the form.
 In other words, $\nu$ has dimensions of only mass and is defined as a function of both a point and a volume element.
-In contrast, the density $\rho$ is just a function of the coordinates at a point, not the volume element, and has dimensions of mass over volume.
+In contrast, the density $\rho$ is just a function of the coordinates at a point, not the volume element, and has dimensions of mass over units.
 This shows that the coordinate-independent function $\nu$ cannot play the role of the coordinate-dependent density $\rho$.
@@
 a_n}$ is the Levi-Civita symbol.
-We can then express any differential density $\nu = \hat{\rho} dV$ as the product of the volume density times a scalar function $\hat{\rho} \equiv \rho \ \epsilon \sqrt{|g_{ab}|}$.
-At first glance, this expression misleadingly suggests that the density $\hat{\rho}$ is coordinate-independent: it is a scalar quantity with the correct dimensions of mass over volume.
-However, on physical grounds, we require that the numerical value of $\hat{\rho} $ equals the value of $\rho$.
-This is only possible if the determinant of $g_{ab}$ equals one.
-Hence, we must define $\hat{\rho} $ in a special coordinate system, namely one where $|g_{ab}| = 1$.
-For this reason, $\hat{\rho} $ remains coordinate-dependent.
+This gives us an objective way to assign volumes to regions of space.
+We can then express any differential volume form $\nu = \hat{\rho} dV$ as the product of the volume form times a scalar function $\hat{\rho} \equiv \rho \ \epsilon \sqrt{|g_{ab}|}$.
+At first glance, this expression misleadingly suggests that $\hat{\rho}$ is a coordinate-independent density: it is a scalar quantity with the correct dimensions of mass over volume, meeting our standard criteria for a density.
+However, note that the numerical value of $\hat{\rho} $ equals the value of $\rho$ only if $|g_{ab}| = 1$, where $1$ is a pure number.
+This match happens only in special coordinate frames and, for this reason, $\hat{\rho} $ remains coordinate-dependent.
+In other words, the metric tensor defines a unit for the volume form (e.g.
+meters cubed), and we always express the volume in those units, regardless of what coordinates we picked.
+That would be the natural choice of volume in a Cartesian system, for example, but not in a spherical one.
+%That is: to This is only possible if the determinant of $g_{ab}$ equals one. Hence, we must define $\hat{\rho} $ in a special coordinate system, namely one where $|g_{ab}| = 1$. 
 %det g_ab =1 means cells you construct are equal volume. whereas cartesian means not just equal volume but sides (like cubes)
@@
 
-For example, in Cartesian coordinates $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$.
-Whereas in polar coordinates, $|g_{ab}| = r^2$ and hence the volume form $dV = r dr \wedge d\theta$.
-This means that $\hat{\rho}$ at best expresses the density in a particularly convenient Cartesian-like coordinate system: one in which we can compare the density point-wise.\footnote{This is similar to what happens with the concept of proper time: it is defined in the co-moving frame, so it is coordinate dependent.
-Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate invariant.
+For example, the metric tensor of a two dimensional space may define length in terms of meters.
+The canonical volume form would then define areas in meters squared.
+This would be the natural choice of volume unit in Cartesian coordinates, where in fact $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$.
+In polar coordinates, the natural choice of units would be meter $\cdot$ radians.
+However, the volume form will still return meters squared, since we will have $|g_{ab}| = r^2$ and $dV = r dr \wedge d\theta$.
+This means that $\hat{\rho}$ at best expresses the density in a particularly convenient Cartesian-like coordinate system: one in which the volume is expressed in special units and we can compare the density point-wise.\footnote{This is similar to what happens with the concept of proper time: it is defined in the co-moving frame, so it is coordinate-dependent.
+Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate-invariant.
 This coordinate-invariance derives from the fact that we have conventionally agreed to define proper time in terms of the object's co-moving frame, rather than in terms of a different frame.} By conventionally agreeing to work in a coordinate system where $|g_{ab}| = 1$, we can make the density $\hat{\rho}$ coordinate-invariant.
@@
 In the case of massive particles, where the Hamiltonian (and the Lagrangian) are quadratic, the Hamiltonian of Equation~\ref{massive_Hamiltonian} entails that $\frac{\partial H}{\partial p_k \partial p_j} = m g^{kj}$.
-This is the case where the densities in the tangent bundle can be properly understood in terms of Riemmanian geometry.
+This is the case where the densities in the tangent bundle can be properly understood in terms of Riemannian geometry.
 For the more general case, we note that
@@
 
-%\subsection{Fundamental density and gauge-redundant momentum}
+%\subsection{Fundamental density and gauge-dependent momentum}
 \subsection{Coordinate-dependent yet fundamental} %; Coordinate-variant yet significant
@@
 % A defender of the necessity of coordinate-independence for fundamentality might respond as follows:
+
 % % [To do: provide literature references for this and motivating quotations; likewise for earlier remark about intrinsicality: Field, North, Chen]
@@
 Being necessary for fundamental significance does not, however, make coordinate-invariance necessary for \textit{non-fundamental} physical significance.
-Some quantities and properties are coordinate-variant yet still physically significant.\footnote{Indeed, many physically significant quantities are gauge-redundant, such as conjugate momentum and the Lagrangian (see Sections~\ref{gauge_momentum} and~\ref{gauge}).
-Insofar as gauge-redundancy is analogous to being coordinate-variant, this supplies a further reason for denying that coordinate-invariance is a necessary condition for physical significance.} Our position here again places us at odds with others in the literature.
+[XXX: disagree] Some quantities and properties are coordinate-variant yet still physically significant.\footnote{Indeed, many physically significant quantities are gauge-dependent, such as conjugate momentum and the Lagrangian (see Sections~\ref{gauge_momentum} and~\ref{gauge}).
+Insofar as gauge-dependency is analogous to being coordinate-variant, this supplies a further reason for denying that coordinate-invariance is a necessary condition for physical significance.} Our position here again places us at odds with others in the literature.
 For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
-However, some choices of unit-variables $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [XXX to do: provide a clear example of this].
+However, some choices of unit-variables $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [XXX to do: provide a clear example of this ---  not sure we can make this claim].
 Similarly, we saw in Section~\ref{dependence} that some coordinate choices in Lagrangian mechanics are physically significant because they enable the property of point-wise comparison.
@@
 But they're not necessary for being physically significant.
-Both coordinate-variant and gauge-redundant quantities can still be physically significant, although they cannot be physically fundamental.
-So in this way, we view gauge-redundancy as a kind of coordinate-variants (rather than a kind of coordinate-dependence).
+Both coordinate-variant and gauge-dependent quantities can still be physically significant, although they cannot be physically fundamental.
+So in this way, we view gauge-dependency as a kind of coordinate-variance (rather than a kind of coordinate-dependence).
 And gauge-invariance is analogous to a kind of coordinate-invariance, rather than a kind of coordinate-independence.
@@
 % Question: is any quantity that is coordinate-variant automatically coordinate-dependent?
-If so, then coordinate-variances a special case of coordinate-dependence?
+If so, then is coordinate-variance a special case of coordinate-dependence?
 And so gauge-variance would also be a special case of coordinate-dependence?
@@
 
-%The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density (as the fundamental description of a classical state)?
+%The gauge-dependency of the conjugate momentum supplies further grounds for privileging the invariant phase space density (as the fundamental description of a classical state)?
 
@@
 Likewise, \textcites[164,171]{Wallace} claims that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
-Yet, as we now show, the conjugate momentum $p$ is actually gauge-redundant.
+Yet, as we now show, the conjugate momentum $p$ is actually gauge-dependent.
 Assuming that gauge-invariance is a necessary condition for fundamental physical significance, this precludes $p$ from having fundamental status.
@@
 Both the invariant density and the canonical state variable pair are fundamental in the physicist sense.
-But the invariant density is more fundamental because it does not involve gauge-redundancy.
-
-To show that the conjugate momentum is gauge-redundant, consider the set of transformations that leave the density, position, and velocity invariant.
+But the invariant density is more fundamental because it is not gauge-dependent.
+
+Before proceeding, we want to give a precise account of what gauge redundancy physically represents.
+In the spirit of our previous constructions, we want to find a physically-motivated definition that leads to the correct mathematical structure.
+We first note that, when motion equivalence holds, we can completely characterize states in terms of the position $q$ and its first derivative $\dot{q}$.
+Hence, any canonical transformation in phase space that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant.
+We will show that these transformations are exactly the gauge transformations.
+
+% To show that the conjugate momentum is gauge-dependent, consider the set of transformations that leave the density, position, and velocity invariant.
 We will see that we can interpret these transformations as a set of gauge transformations on the conjugate momentum $p $.
@@
 We will further assume that the transformations leave the density invariant, which entails that the new variables must be canonical.
-Hence, the new state variables $(q, \hat{p})$ must respect the usual Poisson brackets:
+
+Since the position does not change, the new state variables can be written as $(q, \hat{p})$. Since the transformation is canonical, the new state variables must respect the usual Poisson brackets:
 \begin{equation}
@@
 Therefore, changes of state variable of the form $\hat{p}_j = p_j + \partial_j f$ do not change either the equations of motion or the trajectories.
-From the usual notion of a gauge transformation, we recognize these as a family of gauge transformations that change the conjugate momentum variable $p$ while leaving $(q,\dot{q}) $ invariant.
-We can thereby physically interpret this set of gauge transformations as the transformations that leave the density and $(q,\dot{q}) $ invariant.
-
-The gauge-redundancy of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
+We recognize these as the usual family of gauge transformations.
+We can thereby claim that, in the context of classical particle mechanics, the gauge transformations are the state variable transformations that leave the density and $(q,\dot{q}) $ invariant.
+
+The gauge-dependency of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
 Insofar as coordinate-invariance is a plausible necessary condition for fundamental physical status, so is gauge-invariance.
-By being gauge-variant, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can remain physically significant but not fundamental).
-
-Additionally, the gauge redundancy of conjugate momentum shows that not all state variables are on equal physical footing.
+[XXX: check consistency with previous claim] By being gauge-variant, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can remain physically significant but not fundamental).
+
+Additionally, the gauge-dependency of conjugate momentum shows that not all state variables are on equal physical footing.
 Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.
@@
 % % [To do: think about whether treating gauge-invariance as a necessary condition for fundamentality is in tension with our position, specifically our contention that coordinate-dependency does NOT preclude a quantity from fundamental physical significance.
-Does our position require an important physical difference between gauge-redundancy and coordinate-dependence?].
+Does our position require an important physical difference between gauge-dependency and coordinate-dependence?].
 Maybe the thought is that gauge-invariance and coordinate-invariance are on the same footing.
 So if coordinate-invariance is a necessary condition for physical significance, then so is gauge-invariance.
-A gauge-redundant quantity can still be physically significant, in the same way that a coordinate -dependent quantity can be physically significant.
-% % Note that if coherent, our position has the following consequence: gauge-redundancy is a sufficient condition for a quantity to be non-fundamental (in the same way that coordinate-variance is sufficient for non-fundamentality).
+A gauge-dependent quantity can still be physically significant, in the same way that a coordinate -dependent quantity can be physically significant.
+% % Note that if coherent, our position has the following consequence: gauge-dependency is a sufficient condition for a quantity to be non-fundamental (in the same way that coordinate-variance is sufficient for non-fundamentality).
 But coordinate-dependency is NOT a sufficient condition for being non-fundamental.
 This could strike some readers as in tension.
-If we are comfortable viewing gauge-redundancy as precluding a quantity from fundamental significance, then why not coordinate-dependence?
-Response: we just need to motivate that gauge-redundancy is more analogous to coordinate-variance than coordinate-dependence.
+If we are comfortable viewing gauge-dependency as precluding a quantity from fundamental significance, then why not coordinate-dependence?
+Response: we just need to motivate that gauge-dependency is more analogous to coordinate-variance than coordinate-dependence.
 
@@
 We show that the action principle holds even when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
-Section~\ref{gauge} discusses how the Lagrangian is highly gauge-redundant, and therefore less physical than the Hamiltonian.
+Section~\ref{gauge} discusses how the Lagrangian is gauge-dependent, and therefore less physical than the Hamiltonian.
 This provides a further argument for privileging Hamiltonian mechanics in the interpretation of classical systems.
@@
 Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition.
-We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating the generalization, see \textcites[110]{Arnold}.
-[XXX To do: verify that Arnold can legitimately be read as generalizing the argument below.]}
+We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating this generalization, see \textcites[236--37, 243--45]{Arnold}.}
+%our derivation seems very similar to Arnold chapter 9, section 45:C: ``the principle of least action in phase space''. Note that we briefly discussed Arnold on 11-23-20 meeting, around 12:30. 
 % Note on the generalization: cannot use curl.
@@
 
-Nonetheless, reformulating Hamiltonian mechanics within the extended phase space provides one way to rectify these deficiencies.
+Nonetheless, reformulating Hamiltonian mechanics within the extended phase space $(q_i, p_i, t, E)$ provides one way to rectify these deficiencies.
 At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and an affine parameter within the usual Hamiltonian framework.
@@
 Next, we define the phase space evolution over an affine parameter $s $.
-In this extended Hamiltonian framework, the Hamiltonian constraint functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [XXX to do: clarify what the `Hamiltonian constraint' is].
-Crucially, the Hamiltonian constraint is now coordinate-invariant, even with respect to time-dependent coordinate transformations.
-For example, if we define the affine parameter as the proper time, then the Hamiltonian constraint is proportional to the mass (which is manifestly coordinate-invariant).
-This formalism thereby provides a clear physical meaning for the Hamiltonian constraint.
-[XXX To do: provide a reference for extened phase space formulation].
+This yields two more equations of motion: $dt/ds = -\partial \mathcal{H}/ \partial E$ and $dE/ds = -\partial \mathcal{H}/ \partial t$, where $\mathcal{H}(q_i, p_i, t, E)$ is the extended Hamiltonian.
+In this framework, there is a \textit{Hamiltonian constraint} $\mathcal{H}=0=H(q_i, p_i, t)-E$ that functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [XXX to do: provide a reference for extened phase space formulation.
+And check this discussion of the `Hamiltonian constraint'].
+Crucially, the extended Hamiltonian---unlike the ordinary Hamiltonian---is coordinate-invariant, even with respect to time-dependent coordinate transformations.
+For example, if we define the affine parameter as the proper time, then the extended Hamiltonian is proportional to the mass (which is manifestly coordinate-invariant).
+This provides a clear physical meaning for the extended Hamiltonian.
 We have now considered many conceptual and physical advantages of the Hamiltonian formalism.
 In the next section, we rebut Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.
+%%Arnold uses a related extended phase space $(q_i, p_i, t)$ in his chapter 9, but doesn't provide a general discussion of it. 
+%Discussion by GC of extended phase space: https://tinyurl.com/4tlbej76
 % Is the Hamiltonian constraint for the extended phase space the same as the discussion in this Wikipedia page: https://en.wikipedia.org/wiki/Hamiltonian_constraint
@@
 There, we saw how both the linear relationship between momentum and velocity and the quadratic form (\ref{6.2}) follow from imposing the motion equivalence condition.
-In Equation~\ref{linear_relationship}, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (connected to the gauge-redundancy of the conjugate momentum and the electromagnetic field).
+In Equation~\ref{linear_relationship}, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (connected to the gauge-dependency of the conjugate momentum and the electromagnetic field).
 Just as Curiel does, we then used this relationship to derive a quadratic form for the Hamiltonian (Equation~\ref{massive_Hamiltonian}).
@@
 His definition takes satisfying certain kinematical constraints as fundamental.
-[XXX: state defined at equal time, velocity is not] In contrast, we privilege a system having a unique solution, i.e.
+[the state is defined at equal time; the velocity is not] In contrast, we privilege a system having a unique solution, i.e.
 being deterministic and reversible.
@@
 Our formalism emphasizes the fundamental significance of the invariant phase space density.
-Whereas the state variables $(q, p)$ suffer from the gauge-redundancy of the conjugate momentum, the density provides an invariant description within the Hamiltonian formalism.
+Whereas the state variables $(q, p)$ suffer from the gauge-dependency of the conjugate momentum, the density provides an invariant description within the Hamiltonian formalism.
 This explains why statistical mechanics is best done on phase space, since it provides the natural setting for an invariant density.

### 2021-02-22 Overviews/Experimental verifiability/ExperimentalVerifiabilityOverview.tex
Starting overview of experimental verifiability

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-1}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\begin{document}

\title{Assumptions of Physics overview: \\
	experimental verifiability and topology }
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
We briefly show how the use of topological spaces and $\sigma$-algebras in physics can be rederived and understood as the fundamental requirement of experimental verifiability. A set of experimentally distinguishable objects will necessarily be endowed with a topology that is Kolmogorov (i.e. $T_0$) and second countable, which puts severe constrains on the space of well-formed scientific theories. These idea can be taken as a first step in a general mathematical theory for experimental science. This work is an overview of some of the results of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
\end{abstract}

\maketitle

\section{Introduction}

The overall structure (see \cite{AoPBook} for more details) can be summed up in the following diagram that can be used as a guide throughout this note.

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{Diagram.png}
\end{figure}

The set of points of a topological space represent the possible cases (e.g. all possible values for the mass of a particle). Each element in the topology (i.e. each open set) represents a statement that can be experimentally verified (e.g. the mass of the particle is within a finite precision interval). Each element in the Borel algebra represent a statement that, though it may or may not be experimentally verified, is sound within the theory (e.g. the mass of the particle is exactly zero, the mass of the particle is an irrational number). The Assumptions of Physics framework rederives this structure from the idea that a scientific theory must be logically consistent and evidence based.

\section{Logical context and fundamental axioms}

\section{Experimental and theoretical domains}

\section{Topologies and $\sigma$-algebras}

\section{Experimental relationships and continuous functions}

\bibliography{bibliography}

\end{document}

### 2021-02-22 Articles/2019-Entropy/Source/notes.tex
Images and old preliminary notes for thermodynamics

\documentclass{article}
\usepackage{authblk}

\title{Title}
\author[1]{Gabriele Carcassi}
\author[1]{Christine A. Aidala}
\affil[1]{
	Physics Department\\
	University of Michigan \\
	Ann Arbor, MI 48109
}

\date{\today}
\begin{document}

%	\maketitle

	\section{Notes when reading Uffink}

\begin{quote}
	Another point to be noted is that Clausius’ result that the entropy in an adiabatically isolated
	system can never decrease is derived from the assumption that one can find a quasistatic process
	that connects the final to the initial state, in order to complete a cycle. Indeed, if such a process
	did not exist, the entropy difference of these two states would not be defined. The existence of
	such quasistatic processes is not problematic in many intended applications (e.g. if sf and si are
	equilibrium states of a fluid); but it may be far from obvious in more general settings (for instance
	if one considers processes far from equilibrium in a complex system, such as a living cell). This
	warning that the increase of entropy is thus conditional on the existence of quasistatic transitions has
	been pointed out already by Kirchhoff (1894, p. 69).
\end{quote}
We require being able to isolate the system.

An important difference is in the notion of reversibility. Thermodynamics reversibility requires the existence of a process that can ``run the system in reverse''. Our starting point, which we may call dynamical reversibility, is simply the ability to reconstruct the initial state given the final, the dual of determinism. The key insight is that a system undergoes a thermodynamically reversible process if and only if the whole (i.e. system plus couple system) undergo a dynamically reversible process.

impinges on the ability that can ``run the system in reverse'' or, at the very least, quasistatic transitions. The second law is often defined in terms of this concept and therefore the very notion of entropy as a state variable relies on this notion (CITE Kirchoff 1894).

In standard thermodynamics, the notion of entropy as a state variable is conditional on the existence of processes that can ``run the system in reverse'' or that, at the very least, quasistatic transitions (CITE something) which is, in many cases, a problematic assumption .  In our work, we d the notion of reversibility is simply the dual of determinism. The  (and therefore indirectly ) is conditional on the existence of a process that can ``undo'' the work. In our framework,

One advantage of this approach is substituting the In standard thermodynamics the

\begin{quote}
	This classical conception of probability would, of course, remain a view without any bite, if it
	were not accompanied by some rule for assigning values to probabilities in specific cases. The only
	such available rule is the so-called ‘principle of insufficient reason’: whenever we have no reason to
	believe that one case rather than another is realized, we should assign them equal probabilities (cf.
	Uffink 1995). A closely related version is the rule that two or more variables should be independent
	whenever we have no reason to believe that they influence each other.
\end{quote}
We get equal probability from equal description.

\begin{quote}
	In this respect, his position stands in sharp contrast to that of Boltzmann, who made the project of	finding this unified basis [between statistical and mechanical] his lifework.
\end{quote}

\begin{quote}
	Modern commentators are utterly divided in the search for a direction in which a motivation
	for the choice of the size of these cells can be found. Some argue that the choice should be made
	in accordance with the actual finite resolution of measuring instruments or human observation capabilities.
	The question whether these do in fact favour a partition into cells of equal phase space
	volume has hardly been touched upon.
\end{quote}
We take invariance of densities as giving the correct measure.

Note somewhere: we are not interested in demonstrating that processes go to equilibria. This does not always happen anyway, so we can't demostrate it in general. What we want to show that, if we have a process that tends to an equilibrium, the entropy is maximized at equilibrium. And that, in a thermodynamic context, that entropy corresponds to the thermodynamic entropy.

\begin{quote}
	He approaches this issue quite cautiously, by pointing out certain analogies
	between relations holding for the canonical and microcanonical ensembles and results of thermodynamics.
	At no point does Gibbs claim to have reduced thermodynamics to statistical mechanics.
\end{quote}
	\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

### 2021-02-22 Articles/2020-MeasureTheoryFoundation/main.tex
Checkpointing article for measure theory

\documentclass{article}
\usepackage[utf8]{inputenc}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
% Remove line spaces between items of enumerate and itemize
\usepackage{enumitem}
\setlist{noitemsep}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\input{basicstyle}
\input{logic}
\input{theorems}

\title{Description granularity: a common foundation for geometry, measures and probability.}
\author{gabriele.carcassi }
\date{September 2020}

\begin{document}

\maketitle

\begin{abstract}
    In this paper we present an order theoretic framework that we hope can be used as a common foundation for geometrical, measure theoretic and probabilistic concepts in the sciences. As in comparative probability, we start with a $\sigma$-complete Boolean algebra that represents all possible descriptions in a given context and impose a preorder. In our case, the preorder does not represent likelyhood comparison, but rather whether one description is at a finer level of granularity than another. At this point, any statement (apart from an impossible one) can be chosen to act as a unit, and a measure (not necessarily bounded) can be constructed to quantify the degree to which the other statements give a more or less refined account. The starting lattice is not necessarily Archimedean, therefore the choice of unit will decide which elements are finite, infinite or infinitesimal. The family of measures so obtained are able to give a size to sets at different level of scale (e.g. points, lines, surfaces, ...) much like geometrical structures do. Our hope is that this structure recovers as special cases the geometrical, measure theoretical and probabilistic constructions we currently use in the sciences while giving a generalization that still works in the case where they fail (e.g. lack of $\sigma$-additivity in quantum probability) or expected to fail (e.g. breakdown at Planck's length).

	This work is part of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
\end{abstract}

\def\eqgran{\doteq}
\def\finer{\leqdot}
\def\nfiner{\nleqdot}
\def\coarser{\geqdot}
\def\ncoarser{\ngeqdot}
\def\sfiner{\lessdot}
\def\scoarser{\gtrdot}

\section{Introduction}

This working draft summarizes the status of our work on a new approach to the foundations of geometry and measure theory and is intended to gather early feedback. The goal is to rederive the usual concepts in geometry, measure theory and probability theory from a simpler more general structure, namely a preorder. This construction will force us to make explicit the physical requirements that are baked into the use of the common concepts and provide us with a framework that is, hopefully, still valid when those assumptions fail.

The general idea is similar to what is done in comparative probability, yet we want the techniques extend from a single measure to a family of measures that allow us to define and compare sizes of sets at different scales (e.g. points, lines, surfaces, ...). We start with a $\sigma$-complete Boolean algebra $\tdomain$, a theoretical domain, that represents a set of statements, a set of descriptions, for a physical system in a given context. A Boolean algebra is also a lattice, and therefore already comes equipped with a partial order $\narrower$ such that $\stmt_1 \narrower \stmt_2$ if $\stmt_1 \OR \stmt_2 \equiv \stmt_2$. This partial order indicates whether one statement gives a \textbf{narrower}, more specific, description than the other. For example:
\begin{itemize}
    \item \statement{The position of the object is between 0 and 1 meters} $\narrower$ \statement{The position of the object is between 0 and 1 kilometers}
    \item \statement{The fair die landed on 1} $\narrower$ \statement{The fair die landed on 1 or 2}
    \item \statement{The first bit is 0 and the second bit is 1} $\narrower$ \statement{The first bit is 0}
\end{itemize}
In these cases, the first statement has to be ``contained'' in the second, which is more general.

We then suppose we have a way to compare any two statements and decide which one provides a description at a finer level of granularity. This defines an additional preorder $\finer : \tdomain \times \tdomain \to \mathbb{B}$. Saying $\stmt_1 \finer \stmt_2$ means that the description provided by $\stmt_1$ is \textbf{finer}, gives more information, is more precise, than the description provided by $\stmt_2$. For example:
\begin{itemize}
    \item \statement{The position of the object is between 0 and 1 meters} $\finer$ \statement{The position of the object is between 2 and 3 kilometers}
    \item \statement{The fair die landed on 1} $\finer$ \statement{The fair die landed on 3 or 4}
    \item \statement{The first bit is 0 and the second bit is 1} $\finer$ \statement{The third bit is 0}
\end{itemize}
In these cases, the first statement may not be contained or overlap with the second. Even though the statements come from different disciplines (i.e. geometry, probability and information theory) they all work nicely with the same notion of fineness.

Note that, conceptually, a statement is finer than another if it is true in fewer cases. Therefore the notion of fineness is linked to the notion of ``counting'' of the possible outcomes. It should be no surprise, then, that the same concept is compatible with the different disciplines above: geometrical size quantifies the ways we can arrange objects; probability theory quantifies the ratio of expected cases; information theory needs to quantify the different possible messages stored, computed or transmitted through an information system. Therefore the common structure we propose is not simply a mathematical pattern among disconnected structures: it is a single conceptual framework.

While we borrow many ideas from comparative probability, our goal cannot be achieved with a single measure. A single measure can only compare objects with a finite measure. All objects with zero measure (or infinite measure) are indistinguishable. For example, we want to say:
\begin{itemize}
    \item $\stmt_1$ = \statement{The horizontal position of the object is exactly 0 meters}
    \item $\stmt_2$ = \statement{The horizontal position of the object is exactly 1 or 2 meters}
    \item $\stmt_3$ = \statement{The horizontal position of the object is between 0.5 and 1.5 meters}
    \item $\stmt_4$ = \statement{The horizontal position of the object is between 1.5 and 3.5 meters}
    \item $\stmt_1 \finer \stmt_2 \finer \stmt_3 \finer \stmt_4$
    \item $\stmt_1 \ncoarser \stmt_2 \ncoarser \stmt_3 \ncoarser \stmt_4$
\end{itemize}
Therefore we need to construct a family of measures, each identified by picking a particular statement that acts as a reference, as a unit.

On the other hand, geometric structures do provide measure for different dimensionality by starting from a lowest definition (i.e. line distance for Riemannian and surface are for symplectic) and then constructing higher dimensional measures through products. While want to recover those cases as well, we want to be able to capture the physical idea of a unit system, which is not typically done. In that case, there will be dimensions that are not directly comparable (e.g. pressure, volume, number of particles, chemical potential) but are still comparable under appropriate composition (i.e. pressure times volume with number of particles times chemical potential).

\section{Overview}

In this overview we give a more detailed account of our goal and the status of our search. This will give a general idea of the mathematical issues, but it will also focus on the construction at a physical/conceptual level. It is important for our larger project that the mathematical detail proceeds in lock step with the conceptual and physical understanding. Later sections, instead, will concentrate on the bare formal aspects.

\subsection{Fineness and granularity}

We start with a $\sigma$-complete Boolean algebra $\tdomain$, which we call \textbf{theoretical domain}, that can be generated from a countable subset, a countable basis. Conceptually, this represents all the statements that can be associated with an experimental verification procedure. The full detail of why this is the correct abstraction is discussed in detail in our previous work.\footnote{See \url{https://assumptionsofphysics.org/book/}} It is a property of this type of spaces that each statement $\stmt \in \tdomain$ can be expressed as a disjunction of possible cases, of the  \textbf{possibilities} of the domain. Formally, each possibility is a minterm\footnote{A minterm is a disjunction in which each term appears once, either negated or not.} of the countable basis. The theoretical domain can then equivalently be represented by a $\sigma$-algebra.\footnote{This conceptually maps perfectly to what one has in probability: a sample space (i.e. the set of possibilities) and a set of events (i.e. the statements in the domain). The advantage of our ``pointless'' constructions is that we can treat domains, subdomains and composite domains in the same manner.}

A Boolean algebra is also a lattice, and therefore $\tdomain$ already comes with a partial order $\narrower$ such that $\stmt_1 \narrower \stmt_2$ if $\stmt_1 \OR \stmt_2 \equiv \stmt_2$. We call \textbf{narrowness} this partial order as it describes whether a statement gives a narrower description of another as we saw in the introduction. The narrowest statement in the domain is the impossibility $\impossibility$ and the broadest one is the certainty $\certainty$.\footnote{In terms of the $\sigma$-algebra, narrowness corresponds to set inclusion, the certainty is the full set and the impossibility is the empty set.}

Conceptually, narrowness allows us to say that a description is more refined than another only if the first is fully contained in the second. It cannot compare descriptions that do not overlap, that are incompatible. That is why it is not sufficient to quantify the granularity, the level of detail, of the description. The ability to compare the granularity of the description of two statements, therefore, is captured by an additional binary relationship $\finer$ we call \textbf{fineness}.\footnote{In terms of the $\sigma$-algebra, fineness allows us to say that one set is ``bigger'' than another.}

The whole program can be summarized in the following question: what intuitive conceptual properties does fineness have physically that, once properly formalized mathematically, allow us to recover measures, probability and geometrical structures?\footnote{In terms of the $\sigma$-algebra, how do we go from being able to compare the size of sets to being able to quantify it?}

\subsubsection{Necessary properties for fineness}

The first set of properties are the ones we must expect fineness to have in all cases. At this point, there are three, identified by DeFinetti for comparative probability, that we believe are necessary and that work as simple and intuitive starting points.

\textbf{Boundness}. As we mentioned in the introduction, a description is finer than another if it can be true in fewer cases. This intuitive notion is in general ill-defined except for two cases. A certainty $\certainty$ (e.g. \statement{the position is between minus infinity and plus infinity}) is true in all cases, therefore it must be the coarsest statement. An impossibility (e.g. \statement{the position is between 0 and 1 and also between 2 and 3}) is false in all cases, therefore it must be the finest statement. This means that fineness must be uniquely bound by the impossibility and the certainty. Any contingent statement must be strictly between the two since there are some cases in which it will be true and others in which it will be false. This is the first required property.\footnote{In terms of the $\sigma$-algebra, the full set is strictly the biggest set and the empty set is strictly the smallest set.}

\textbf{Transitivity}. Additionally, if $\stmt_1$ gives finer description than $\stmt_2$ and $\stmt_2$ gives a finer description than $\stmt_3$ then it must be that $\stmt_1$ gives a finer description than $\stmt_3$. This means that fineness is transitive, and this is the second required property.

\textbf{Offset monotonicity}. Lastly, if we take two descriptions and we make them less precise ``in the same way'', we must have the same relationship in fineness. More precisely, suppose we have two descriptions $\stmt_1$ and $\stmt_2$ and we pick a third description $\stmt$ that is incompatible\footnote{Two statements are incompatible if they can't be both true at the same time: their conjunction $\stmt_1 \AND \stmt \equiv \impossibility$ is impossible. This is the analogue of disjoint sets in a $\sigma$-algebra.} with both. The respective disjunctions $\stmt_1 \OR \stmt$ and $\stmt_2 \OR \stmt$ will be respectively ``equally'' broader than the respective original statement. Therefore we will want them to be ``equally'' coarser as well. In other words, it must be that $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \OR \stmt \finer \stmt_2 \OR \stmt$: fineness must be offset monotonic\footnote{Some authors call this property ``monotony'', which we don't find descriptive. We prefer ``offset monotonicity'' as it makes it clear that adding or removing a ``constant'' on both sides is a monotonic operation.} in the sense that it is invariant if we add or remove a statement from both sides, and this is the third and last required property.\footnote{In terms of the $\sigma$-algebra, a set is smaller than another if and only if it is also smaller when we increase both sets with another disjoint from the two.}

From these three properties, we can show the following:
\begin{description}
    \item[Preorder.] Fineness is a preorder over the theoretical domain $\tdomain$.
    \item[Monotonicity.] Fineness respects narrowness: $\stmt_1 \narrower \stmt_2$ implies $\stmt_1 \finer \stmt_2$.
    \item[Equigranularity is an equivalence.] Equigranularity, the symmetrization of fineness, is an equivalence class, and fineness is an order of those equivalence classes. That is, fineness orders the levels of description.
\end{description}

While we are not convinced these are sufficient properties to define fineness, we are not aware of other properties that would make the cut.

\subsubsection{Optional properties for fineness}

Optional properties for fineness include properties that are desirable, but are known to fail in some cases. These will be properties of a particular domain, rather than fineness itself.

\textbf{Equigranularity of possibilities}. Possibilities provide a complete characterization of a domain: if we know one possibility to be true (e.g. \statement{the mass of the photon is exactly 0 eV}), we know the truth value of all other statements (e.g. \statement{the mass of the photon is between x and y eV}). Therefore it is natural to assume that all possibilities are equigranular. In fact, this is a necessary property if we want to have a unique measure when we take a possibility as a unit.\footnote{The measure, in this case, is simply the counting measure and equigranularity of the possibilities plays a role similar to the principle of indifference in classical probability} However, this property can be easily violated by creating a subdomain.

Suppose we have a uniform domain with three possibilities: red, green, blue. We can take a ``color blind'' subdomain that cannot distinguish between red and green. This would have two possibilities: red/green and blue. This will not be uniform. The idea is that if the domain is not uniform, the cases are not really describing ``the same thing'': some information is missing. Equigranularity of possibilities, then, is not a property of fineness itself, but a property of a domain that is ``well-posed''.

\textbf{Totality}. If our objective is to construct a measure, totality is needed in the sense that every statement (i.e. every Borel set of possibilities) needs to be comparable with every other if we are to assign a unique number. However, statements whose size is defined in units of different physical dimensions (i.e. length, pressure, mass) should not be comparable. Lack of totality, in fact, should be a feature that allows us to formally construct a unit system.

Another counter example come from infinite ranges. Take the real line and consider the set of all positive numbers and the set of all negative numbers: should they be comparable? If we say they are equal, then that would tell us that there are more cases before 1 than after. This means that the space is not symmetrical under translation. If the real values represent distance between points, this is bad as it violates the idea that space is uniform. If the real values represents charge, however, this would actually make sense: the idea that there are exactly the same values for positive and negative charges does make sense, and the symmetry we are imposing is natural since zero is really a special number. Therefore different choices of comparability for the same Borel algebra is actually a feature we do not want to lose in general.

However, once we pick the unit, we should be able to find the ``right'' set of statements on which to create the measure. One approach would be to ask the ``largest set of statements mutually comparable in fineness that contain the unit''. In general, we have no expectation that this would be unique. So one question would be: when is this true?

\subsubsection{Unsettled properties for fineness}

\textbf{Boolean lattice}. Since equigranularity classes form a bounded partially ordered set, one can ask whether they should form a lattice. Physically this would make sense: we would be saying that given a set of statements $S \subseteq \tdomain$, we can always find a level of description that is the finest among all coarser descriptions than all the statements (i.e. $\bigveedot S$) and a level of description that is the coarsest among all finest descriptions than all the statements (i.e. $\bigwedgedot S$). However, it is not clear whether this is required, something only to be imposed in some cases or something that can be already proven.

Regardless, the map from a statement to its level of description (i.e. equivalence class in equigranularity), while it is an order isomorphism and preserves the top and bottom, is not a lattice isomorphism. That is, it does not preserve the meet and join. Let us note $\dot{\stmt} = {\stmt}_{/\eqgran}$ the level of granularity of $\stmt$. Now suppose that $\stmt_1 \eqgran \stmt_2 \nequiv \impossibility$ and $\stmt_1 \ncomp \stmt_2$. Then $\dot{\stmt}_1 = \dot{\stmt}_2 \neq \dot{\impossibility}$. Now suppose that $\stmt \to \dot{\stmt}$ is a lattice isomorphism. Then $\dot{\impossibility} = (\stmt_1 \AND \stmt_2)_{/\eqgran} = \dot{\stmt}_1 \wedgedot \dot{\stmt}_2 = \dot{\stmt}_1 \wedgedot \dot{\stmt}_1 = \dot{\stmt}_1$, which contradicts the premise.

\textbf{Addition}. We also need to understand how to define the addition and subtraction. Conceptually, the addition of two level of precision $\dot{\stmt}_1 + \dot{\stmt}_2$ means returning the level of precision that can be achieved by taking the disjunction $\stmt_1 \OR \stmt_2$ of two incompatible statements $\stmt_1 \in \dot{\stmt}_1$ and $\stmt_2 \in \dot{\stmt}_2$. This, of course, is not possible if the statements are ``bigger'' than ``half'' the space. A more order theoretical definition would be to take all possible disjuctions of $\stmt_1 \OR \stmt_2$ such that $\stmt_1 \in \dot{\stmt}_1$ and $\stmt_2 \in \dot{\stmt}_2$ and return the coarsest level of description. That is, the sum is the least upper bound in fineness of all possible disjunctions. One should be able to prove that the incompatible disjunction corresponds to that level of description.

It is not clear whether the axioms above are enough to show that the addition always exists. It may be the existence of the addition, or a premise that leads to it, needs to be added. It is not clear whether, physically, such addition always makes sense, particularly in the context of incomparable descriptions (e.g. descriptions quantified in different units).

\textbf{Product}. Another thing that needs to be defined/recovered is that independent domains lead to the product of the measure. This means that there must be some compatibility condition between statement independence and fineness. It is not clear whether this compatibility is already embedded in the given axioms or not. The issue is that, in comparative probability, independence is only defined later in terms of the probability measure. In our framework, we have already a notion of independence that is defined on statements and logic operation, which is based on the idea of whether fixing the truth value of some statements will also fix the truth value of other statements.

\subsection{A family of measures}

The main goal is to understand under what conditions we can pick a statement $\stmt[u]$ and find a unique measure $\mu_{\stmt[u]} : \tdomain \to \mathbb{R}$ with the following properties:
\begin{description}
    \item[Normalization.] The value for the unit is one,  $\mu_{\stmt[u]}(\stmt[u]) = 1$.
    \item[Monotonicity.] It is a monotonic function with respect to fineness; if $\stmt_1 \finer \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
    \item[Finite additivity.] If $\stmt_1 \ncomp \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$.
\end{description}

\textbf{Countable additivity}. Additionally, we want to understand when finite additivity can be extended to countable additivity. In general, countable additivity will fail. Consider a space with countably many possibilities, all of them equigranular. If we take the certainty as a unit, each possibility is infinitely smaller, and therefore it must be assigned a measure zero. Therefore, the sum (i.e. the disjunction) of infinitely many measure zero statements is one.

\textbf{Equigranular possibilities}. If a possibility is taken as a unit, then the equigranularity of possibilities is a sufficient condition and it leads to the counting measure on the possibilities. This is strictly not necessary: we could have a domain with 5 possibilities $\{x_1, x_2, x_3, y_1, y_2\}$ such that $x_1 \eqgran x_2 \eqgran x_3$, $y_1 \eqgran y_2$ and $x_1 \OR x_2 \OR x_3 \eqgran y_1 \OR y_2$. If we took $x_1$ as a unit, this would lead to a unique measure such that $\mu(x_1)=\mu(x_2)=\mu(x_3)=1$ and $\mu(y_1)=\mu(y_2)=3/2$. This also tells us that we may not be really interested in necessary conditions, since these cases are contrived. Rather, we are interested in sufficient conditions that can be well justified from the physics.

\textbf{Lack of reverse monotonicity}. Monotonicity only works in one direction. If a statement describing an area is taken as unit, all statements about lines, points and countable sets of both are going to be measure zero. Therefore equal measure does not mean equal fineness. This is why fineness can generate a family of measures: it has more information than a single one. When we pick a unit, then, the measure has to ``collapse`` the equivalence classes of equigranularity appropriately. This is tightly linked with a notion of infinitesimal (i.e. the ordering over the level of granularity is non-archemedian, while the real numbers are). The idea is to proceed in the following way:
\begin{enumerate}
    \item pick a unit $\stmt[u]$ and define the notion of infinitesimal
    \item define the preorder $\lequ$ and the equivalence $\equ$ that compares levels of description up to infinitesimals
    \item show that $\lequ$ is order isomorphic to the reals
    \item show that we can pick a particular isomorphism that respects linear addittivity.
\end{enumerate}

\subsubsection{Infinitesimals and finite comparison}

The Archemedian property is typically discussed in terms of linearly ordered groups. \footnote{See \url{https://en.wikipedia.org/wiki/Archimedean_property} for an introduction.} Roughly speaking, if we pick an element $y$ an element $x$ is infinitesimal with respect to $y$ if $\sum\limits_{i=1}^n x < y$ for any $n$. The group is Archemedian if no element is infinitesimal with respect to the other. We need to extend this notion to our ordering in fineness.

The first issue, then, is to find the correct definition to extend this concept on ordering. The addition discussed about (i.e. roughly the disjoint union of statements) should play the role of the group operation. Conceptually, a statement $\stmt$ is infinitesimal with respect to another $\stmt[u]$ if for any $n$ we can find a finite set of equigranular incompatible statements whose disjunction is finer than $\stmt[u]$. That is, we need infinitely many statements equigranular to $\stmt$ to cover a statement of the same granularity than $\stmt[u]$.

The main issue here is that, since the ordering is potentially not total, it is not clear how one can prove the existence of said sets of statements even in the simplest cases.

Assuming those issues are solved, infinitesimals come in two types: countable and uncountable. A countable infinitesimal will be such that the disjoint conjunction of countably many equigranular statements will be bigger than the unit. This is the case where, in the end, the measure will only be finitely additive: infinitely many zero measure elements will some to a finite element. If all infinitesimal with respect to the unit are uncountable, instead, countably many zero measure elements will still form a zero measure element.

This insight may give us a way to define ``finite'' and ``infinite'' elements in an absolute way. A finite element has no countable infinitesimal. For example, take the unit interval over the reals and consider the standard Lebesque measure over the Borel sets (which would correspond to our theoretical domain). All sets of measure zero are uncountable infinitesimals with respect to the unit: since the measure is countably additive, any disjoint union of countably many measure zero sets is going to be measure zero, therefore it will not be as big as the unit. Natuarally, if we take an uncountably many sets (i.e. all the singletons) we would cover the whole unit. Conversely, the full real line can be covered by countably many sets of unit size. While the unit interval is infinitesimally small with the respect with the whole line, it is a countable infinitesimal. This definition of ``finiteness'' would map to physical requirements much better than, for example, the topological notion of compactness.

Once a notion of infinitesimals is well defined, we would define a preorder $\lequ$ and an equivalence relationship $\equ$ that compares the level of description of the statements up to infinitesimals.

\subsubsection{Recovering bounded measures}

The next step is to understand how the properties of $\lequ$ would compare to the requirements set by, for example Villegas \cite{villegas}. In the context of comparative probability, there are theorems that allow to go from a preorder to a probability measure. The idea would be to reuse, as much as possible, the ideas from that literature. Once we pick a unit $\stmt[u]$, we can restrict ourselves to all statements narrower than $\stmt[u]$. At that point, we are looking for has all the mathematical properties of a probability measure as the biggest element, $\stmt[u]$, has to have measure one.

One issue will be the problem of countable infinitesimals, which are incompatiable with a countable measure. Therefore one of the assumptions in comparative probability will effectively exclude the case. We need to understand how that works, so that we can adapt the proofs and allow for only finitely additive measures.

The other issue is that the measure will not be, in general, on the whole domain $\tdomain$, but on a subset $\tdomain_{\stmt[u]}$ that must be totally comparable with $\lequ$. It is not clear how to recover this set. It may be that the equivalence classes of $\equ$ will make our job easier: if two infinitesimal are not comparable in fineness, they will still be merged together. Therefore  incomparable statements may be comparable in $\lequ$ if they can be compared up to an infinitesimal.

The goal here is to solve the above problems to be find the prerequisites for the following proposition: given a statement $\stmt[u] \in \tdomain$ we can define a maximal set of mutually comparable statements in $\lequ$ that contains $\stmt[u]$, upon which we can define a unique $\mu_{\stmt[u]} : \tdomain_{\stmt[u]} \to \mathbb{R}$ such that:
\begin{description}
    \item[Normalization.] The value for the unit is one,  $\mu_{\stmt[u]}(\stmt[u]) = 1$.
    \item[Order isomorphism.] $\stmt_1 \lequ \stmt_2$ if and only if $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
    \item[Finite additivity.] If $\stmt_1 \ncomp \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$.
\end{description}
Moreover, if $\stmt[u]$ has no countable infinitesimals, $\mu_{\stmt[u]}$ is countably additive.

\subsubsection{Extending the measures}

Once we know how to construct measures on the sub-algebra of statements narrower than the unit, we should be able to ``stitch them together'' in the following way.

\begin{enumerate}
    \item Prove that if $\stmt[v] \narrower \stmt[u]$ then $\mu_{\stmt[u]}(\stmt[w]) = \mu_{\stmt[u]}(\stmt[v]) \mu_{\stmt[v]}(\stmt[w]) $ for all $\stmt[w] \narrower \stmt[v]$. To achive this, we will need to prove that the infinitesimals with respect to $\stmt[u]$ are exactly the infinitesimals with resepect to $\stmt[v]$.
    \item Given two statements $\stmt[u] \narrower \stmt[v]$, we extend the measure on $\stmt[u]$ by defining $\mu_{\stmt[u]}(\stmt[v]) = 1/ \mu_{\stmt[v]}(\stmt[u])$. If $\mu_{\stmt[v]}(\stmt[u]) = 0$ then $\mu_{\stmt[u]}(\stmt[v]) = \infty$.
    \item To extend the measure on all other statements. Given two statements $\stmt[v]$ and $\stmt[u]$ we can construct the measure $\mu_{\stmt[u]\OR\stmt[v]}$. We have $\mu_{\stmt[u]}(\stmt[v]) = \mu_{\stmt[u]}(\stmt[u]\OR\stmt[v]) \mu_{\stmt[u]\OR\stmt[v]}(\stmt[v])$.
\end{enumerate}

One issue will be, again, proving the consistent of comparability with different units since all the domains $\tdomain_{\stmt[u]}$ do not contain all the statements.

\section{Current attempt}

\begin{defn}
    A \textbf{theoretical domain} $\tdomain$ is a $\sigma$-complete Boolean algebra that can be generated by a countable subset. We call \textbf{narrowness} the partial order of the algebra, that is $x \narrower y$ when $x \OR y = y$. We say $x \in \tdomain$ is a \textbf{possibility} if it is the immediate successor of the impossibility $\impossibility$. That is, $\impossibility \snarrower x$ and there is no $\stmt \in \tdomain$ such that $\impossibility \snarrower \stmt \snarrower x$.
\end{defn}

The fact that the $\sigma$-algebra is generated by a countable set means it is separable. It also means that each element can be constructed as a disjunction of a set of possibilities.

\begin{defn}\label{fineness}
	 Fineness is a binary relationship $\finer : \tdomain \times \tdomain \to \Bool$ with the following properties:
	\begin{itemize}
		\item boundedness: for each contingent statement $\stmt \in \tdomain$ we have $\impossibility \sfiner \stmt \sfiner \certainty$
		\item transitivity: if $\stmt_1 \finer \stmt_2$ and $\stmt_2 \finer \stmt_3$ then $\stmt_1 \finer \stmt_3$
		\item offset monotonicity: if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \finer \stmt_2$ if and only if $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$.
	\end{itemize}
\end{defn}

\begin{prop}\label{fineness_properties}
    Fineness obeys the following properties:
    \begin{enumerate}
        \item if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \sfiner \stmt_2$ if and only if $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$ (these two tell us that monotony can be extend to strict fineness and equigranularity)
        \item if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \eqgran \stmt_2$ if and only if $\stmt \OR \stmt_1 \eqgran \stmt \OR \stmt_2$
        \item if $\stmt_1 \narrower \stmt_2$ then $\stmt_1 \finer \stmt_2$ (these three tell us fineness and narrowness are monotonically ordered)
        \item if $\stmt_1 \snarrower \stmt_2$ then $\stmt_1 \sfiner \stmt_2$
        \item if $\stmt_1 \equiv \stmt_2$ then $\stmt_1 \eqgran \stmt_2$
        \item $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \AND \NOT \stmt_2 \finer \stmt_2 \AND \NOT \stmt_1$ (these three tell us the region of overlap does not matter)
        \item $\stmt_1 \sfiner \stmt_2$ if and only if $\stmt_1 \AND \NOT \stmt_2 \sfiner \stmt_2 \AND \NOT \stmt_1$
        \item $\stmt_1 \eqgran \stmt_2$ if and only if $\stmt_1 \AND \NOT \stmt_2 \eqgran \stmt_2 \AND \NOT \stmt_1$
        \item if $\stmt_1 \finer \stmt_2$ then $\NOT \stmt_2 \finer \NOT \stmt_1$ (these two tell us that fineness works well with negation)
        \item if $\stmt_1 \sfiner \stmt_2$ then $\NOT \stmt_2 \sfiner \NOT \stmt_1$
        \item if $\certainty \finer \stmt$ then $\certainty \narrower \stmt$ and therefore $\stmt \equiv \certainty$ (these two tells us that fineness can recover narrowness for certainties and impossibilities)
        \item if $\stmt \finer \impossibility$ then $\stmt \narrower \impossibility$ and therefore $\stmt \equiv \impossibility$
    \end{enumerate}
\end{prop}
\begin{remark}
Everything is proved without using totality, so that (if we remove it) everything follows.
\end{remark}

\begin{proof}
For 1, let $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$. Suppose $\stmt_1 \sfiner \stmt_2$, then $\stmt_1 \finer \stmt_2$ and $\stmt_1 \ncoarser \stmt_2$. By monotony we have $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \ncoarser \stmt \OR \stmt_2$. Therefore $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$. Conversely, if $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$ then $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \ncoarser \stmt \OR \stmt_2$. By monotony $\stmt_1 \finer \stmt_2$ and $\stmt_1 \ncoarser \stmt_2$, which means $\stmt_1 \sfiner \stmt_2$.

For 2, now suppose $\stmt_1 \eqgran \stmt_2$, then $\stmt_1 \finer \stmt_2$ and $\stmt_1 \coarser \stmt_2$. By monotony we have $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \coarser \stmt \OR \stmt_2$. Therefore $\stmt \OR \stmt_1 \eqgran \stmt \OR \stmt_2$. Conversely, if $\stmt \OR \stmt_1 \eqgran \stmt \OR \stmt_2$ then $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \coarser \stmt \OR \stmt_2$. By monotony $\stmt_1 \finer \stmt_2$ and $\stmt_1 \coarser \stmt_2$, which means $\stmt_1 \eqgran \stmt_2$.

For 3, suppose $\stmt_1 \narrower \stmt_2$. Then $\NOT \stmt_2 \ncomp \stmt_2$ and $\NOT \stmt_2 \ncomp \stmt_1$. Since fineness is uniquely bounded, we have $\NOT \stmt_2 \OR \stmt_1 \finer \certainty$. We have $\certainty \equiv \NOT \stmt_2 \OR \stmt_2 \coarser \NOT \stmt_2 \OR \stmt_1$. By monotony we have $\stmt_2 \coarser \stmt_1$.

For 4, suppose $\stmt_1 \snarrower \stmt_2$. Then $\NOT \stmt_2 \OR \stmt_1$ is a contingent statement and, since fineness is uniquely bounded, $\NOT \stmt_2 \OR \stmt_1 \sfiner \certainty \equiv \NOT \stmt_2 \OR \stmt_2$. By 1 we have $\stmt_1 \sfiner \stmt_2$.

For 5, suppose $\stmt_1 \equiv \stmt_2$. Then $\stmt_1 \narrower \stmt_2$ and $\stmt_2 \narrower \stmt_1$, which means $\stmt_1 \finer \stmt_2$ and $\stmt_1 \finer \stmt_2$ and therefore $\stmt_1 \eqgran \stmt_2$.

For 6, note that $\stmt_1 \equiv (\stmt_1 \AND \NOT \stmt_2) \OR (\stmt_1 \AND \stmt_2)$, $\stmt_2 \equiv (\NOT \stmt_1 \AND \stmt_2) \OR (\stmt_1 \AND \stmt_2)$ with $\stmt_1 \AND \stmt_2 \ncomp \stmt_1 \AND \NOT \stmt_2$ and $\stmt_1 \AND \stmt_2 \ncomp \NOT \stmt_1 \AND \stmt_2$. By monotony, $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \AND \NOT \stmt_2 \finer \NOT \stmt_1 \AND \stmt_2$.

For 7 and 8, use 1 and 2 respectively in the final step.

For 9, note that $\stmt_1 \AND \NOT \stmt_2 \equiv (\NOT \stmt_2) \AND \NOT (\NOT \stmt_1)$ while $\NOT \stmt_1 \AND \stmt_2 \equiv \NOT (\NOT \stmt_2) \AND (\NOT \stmt_1)$. We have $\stmt_1 \finer \stmt_2$ if and only if (by 6)  $\stmt_1 \AND \NOT \stmt_2 \finer \NOT \stmt_1 \AND \stmt_2$ if and only if (by the equivalence) $(\NOT \stmt_2) \AND \NOT (\NOT \stmt_1) \finer \NOT (\NOT \stmt_2) \AND (\NOT \stmt_1)$ if and only if (by 6 again) $\NOT \stmt_2 \finer \NOT \stmt_1$.

For 10, use 7 whenever 6 was used.

For 11, suppose $\certainty \finer \stmt$. By unique boundedness, $\stmt$ can neither be contingent nor impossible. Therefore $\stmt$ must be certain.

For 12, suppose $\stmt \finer \impossibility$. By unique boundedness, $\stmt$ can neither be contingent nor certain. Therefore $\stmt$ must be impossible.

\end{proof}

\begin{prop}
    Fineness is a preorder.
\end{prop}

\begin{proof}
    Let $\stmt \in \tdomain$. Since $\stmt \narrower \stmt$, by \ref{fineness_properties}.3 $\stmt \finer \stmt$ and therefore fineness is reflexive. By \ref{fineness} fineness is also transitive and therefore is a preorder.
\end{proof}

\begin{defn}
    We say $\stmt_1$ is \textbf{equigranular} to $\stmt_2$ (noted $\stmt_1 \eqgran \stmt_2$) if $\stmt_1 \finer \stmt_2$ and $\stmt_2 \finer \stmt_1$.
\end{defn}

\begin{prop}
Other properties useful properties:
\begin{enumerate}
    \item (monotony as in Villegas) if $\stmt[v]_1 \ncomp \stmt[v]_2$, then $\stmt_1 \finer \stmt[v]_1$ and $\stmt_2 \finer \stmt[v]_2$ imply  $\stmt_1 \OR \stmt_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
    \item if $\stmt_1 \ncomp \stmt[v]_1$, $\stmt_2 \ncomp \stmt[v]_2$ and $\stmt[v]_1 \eqgran \stmt[v]_2$, then $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \OR \stmt[v]_1 \finer \stmt_2 \OR \stmt[v]_2$.
\end{enumerate}
\end{prop}

\begin{proof}

\end{proof}

\begin{prop}
    Equigranularity is an equivalence relationship.
\end{prop}

\begin{proof}
    It is an equivalence relationship because it is the symmetrization of a preorder. Since fineness is reflexive then equigranularity is also reflexive: for all $\stmt \in \tdomain$ $\stmt \finer \stmt$ and therefore $\stmt \eqgran \stmt$. Since fineness is transitive then equigranularity is also transitive: $\stmt_1 \eqgran \stmt_2 \eqgran \stmt_3$ means $\stmt_1 \finer \stmt_2 \finer \stmt_3$ therefore $\stmt_1 \finer \stmt_3$ but also $\stmt_3 \finer \stmt_2 \finer \stmt_1$ therefore $\stmt_3 \finer \stmt_1$ and finally $\stmt_1 \eqgran \stmt_3$. Additionally, equigranularity is reflexive by definition and is therefore an equivalence relationship.
\end{proof}

\begin{defn}
    Let $\tdomain_{/\eqgran}$ be the granularity equivalence classes. Let $(\cdot)_{/\eqgran} : \tdomain \to \tdomain_{/\eqgran}$.
\end{defn}

\begin{prop}
    The function $(\cdot)_{/\eqgran}$ is an order homomorphism.
\end{prop}

\begin{proof}
    Let $\stmt_1, \stmt_2 \in \tdomain$. Suppose $\stmt_1 \narrower \stmt_2$ then by \ref{fineness_properties}.3 $\stmt_1 \finer \stmt_2$ and therefore ${\stmt_1}_{/\eqgran} \finer {\stmt_2}_{/\eqgran}$. Then $(\cdot)_{/\eqgran}$ is an order isomorphism.
\end{proof}

\begin{remark}
    Regardless of whether $\tdomain_{/\eqgran}$ is a lattice, $(\cdot)_{/\eqgran}$ can never be a lattice homomorphism. It will be crucial that two equigranular statements can be incompatible, and this in contradiction with $(\cdot)_{/\eqgran}$ being a lattice homomorphism. In fact, suppose $\tdomain_{/\eqgran}$ is a lattice and let $\wedgedot$ be its meet. Suppose that $\stmt_1 \eqgran \stmt_2 \nequiv \impossibility$ and $\stmt_1 \ncomp \stmt_2$. Then ${\stmt_1}_{/\eqgran} = {\stmt_2}_{/\eqgran} \neq \impossibility_{/\eqgran}$. Now suppose that $(\cdot)_{/\eqgran}$ is a homomorphism. Then $\impossibility_{/\eqgran} = {(\stmt_1 \AND \stmt_2)}_{/\eqgran} = {\stmt_1}_{/\eqgran} \wedgedot {\stmt_2}_{/\eqgran} = {\stmt_1}_{/\eqgran} \wedgedot {\stmt_1}_{/\eqgran} = {\stmt_1}_{/\eqgran}$, which contradicts the premise.
\end{remark}

\begin{remark}
We claim there exist domains which have all possibilities equigranular but fineness is not a total order. Suppose we have a theoretical domain in bijection with the power set of the natural numbers, $\mathcal{P}(\mathbb{N})$. The partial order of narrowness may be given by the ordering of the power set by set containment, and the possibilities (which also serve as a countable generating set) are given by $\mathbb{N}$. The hypotheses of the fineness relation (Definition \ref{fineness}) are satisfied if we order contingent statements by cardinality, with the additional requirement that infinite strict subsets of $\mathbb{N}$ are considered strictly finer than the certainty $\mathbb{N}\in\mathcal{P}(\mathbb{N})$. All finite sets will be totally (pre)ordered, but infinite sets will not be ordered. We note that one could just say that $\stmt_1\finer\stmt_2$ for every pair of infinite contingent statements $\stmt_1,\stmt_2$, but even if we remove those from the relation $\finer$, the axioms are still satisfied, and the (pre)order is not total. 
\end{remark}

\begin{defn}
    A theoretical domain $\tdomain$ is \textbf{uniform} if fineness has the following additional properties:
    \begin{itemize}
        \item totality: for all $\stmt_1, \stmt_2 \in \tdomain$ either $\stmt_1 \finer \stmt_2$ or $\stmt_2 \finer \stmt_1$
        \item uniformity: all possibilities are equigranular.
    \end{itemize}

\end{defn}

\begin{defn}
Let $(S,\leq)$ be a poset. A \textbf{chain} is a subset $T\subseteq S$ which is totally ordered (i.e. every pair of elements in $T$ is comparable in $\leq$). A chain is said to be \textbf{maximal} if it is not a proper subset of any other chain, i.e. there are no ``gaps" in the chain. 
\end{defn}
Observe that in a totally ordered set, there is only one maximal chain, namely the entire set.

\begin{defn}
A theoretical domain $\tdomain$ is said to have the \textbf{maximal chains property} (MCP) if given any chain $S\subseteq\tdomain$ totally ordered by narrowness ($\narrower$), the following holds. There exists a maximum chain $S'\supseteq S$, and a maximal chain $C$ in the set of equigranularity equivalence classes, such that for all statements $s \in S'$, there exists an equivalence class $\overline{c}\in C$ with $s\in \overline{c}$.
\end{defn}

\begin{prop}
Suppose there is a (partial) pre-ordering on the countably infinite proper subsets of $\mathbb{N}$ which satisfies offset monotonicity. Then there exists a uniform theoretical domain which lacks the maximal chains property.
\end{prop}
\begin{proof}
Consider the following example. Let the set of possibilities be given by $\mathbb{N}$ and let the theoretical domain be the powerset $\mathcal{P}(\mathbb{N})$. Let fineness be defined as follows. All finite subsets are ordered by their cardinality (with $\impossibility = \emptyset$). Put any ordering on the countably infinite strict subsets of $\mathbb{N}$, and impose that they are strictly coarser than all finite sets and strictly finer than the entire set $\mathbb{N} = \certainty$. This is an example of a uniform theoretical domain, where all possibilities (natural numbers) are equigranular (singleton sets of cardinality 1), and the ordering of equigranular equivalence classes is total.

Consider the following chain $C$ in $\tdomain$: $c_0 = \emptyset$, $c_n = \{1,\ldots, n\}$ for $n<\infty$ and write $c_{\infty} = \mathbb{N}$. Then $c_m \snarrower c_n$ for all $m<n$. We claim that this is a maximal chain (see below for an explanation).

Now, notice that there are no infinite proper subsets of $\mathbb{N}$ which are in the chain $C$. This represents an equigranularity class which is not represented in the maximal chain $C$, which is a failure of the maximal chains condition.

To see why $C$ is maximal, let $c\in\mathcal{P}(\mathbb{N})$ but that $c\neq c_{\lambda}$ for $\lambda\in\mathbb{N}\cup\{0,\infty\}$. Then there exists some $l\in\mathbb{N}$ with $l+1\in c$ but $l\notin c$. Thus $c$ is not broader than $c_p$ for any $p\geq l$, and $c$ is not narrower than $c_q$ for any $q\leq l+1$. It follows that for $c$ to fit in the chain, it must be higher up than $c_{l+1}$ but lower down that $c_l$, which is not possible. 
\end{proof}

\begin{prop}\label{chainimpliesuniform}
Let $\tdomain$ be a theoretical domain with the maximal chains property. Then the following are equivalent:
\begin{enumerate}
    \item the possibilities are comparable
    \item the possibilities are equigranular
    \item all statements compatible with only finitely many possibilities are comparable to each other.
\end{enumerate}
\end{prop}
\begin{proof}

Suppose all possibilities are comparable. Let $p,q$ be possibilities. Consider the two chains $\{\impossibility, p,\certainty\}$ and $\{\impossibility, q, \certainty\}$. By the MCP, these may be extended to maximal chains in narrowness, with corresponding maximal chains in fineness equivalence classes. Then since possibilities are successors of the impossibility, the fineness equivalence classes represent by $p$ and $q$ must both be an immediate successor of the equivalence class representing the impossibility. Now, since all possibilities are comparable, and the fineness class of $p$ is an immediate successor to that of the impossibility, we must have $p\finer q$. Similarly, we must have $q\finer p$. It follows that $p\eqgran q$.

Assume all possibilities are equigranular. We show that any two statements compatible with only finitely many possibilities are comparable. In fact, we will prove that the equigranularity equivalence classes for finite disjunctions of possibilities are given by the size of the disjunction. We proceed by induction on $N$, the larger number of compatible possibilities for the two statements. The case of $N=1$ follows from equigranularity of possibilities. Now suppose the claim holds for some $N$. Let $\stmt_1 = p_1\OR\cdots\OR p_{N+1}$ and $\stmt_2 = q_1\OR\cdots\OR q_m$ be statements compatible with only finitely many possibilities $p_i,i=1,\ldots,N+1$ and $q_j,j=1,\ldots,m$ respectively, where $m\leq N+1$.

Consider first the case where $m\leq N$. Then $\stmt_2 \finer p_1\OR\cdots\OR p_N \sfiner p_1\OR\cdots\OR p_N\OR p_{N+1} = \stmt_1$, so they are comparable, and in particular all statements compatible with fewer than $N+1$ possibilities are strictly finer than $\stmt_1$. Next, suppose $m=N+1$. Write $\stmt_1' = p_1\OR\cdots\OR p_N$ and $\stmt_2' = q_1\OR\cdots\OR q_N$. We have already that all disjunctions of $N$ possibilties are equigranular, so $\stmt_1'\eqgran\stmt_2'$. Possibly by rearranging the order of the possibilties in the disjunction, we may assume $q_{N+1}\ncomp\stmt_1'$ and $q_{N+1}\ncomp\stmt_2'$. By Proposition \ref{fineness_properties}, we have $\stmt_1\OR q_{N+1}\eqgran\stmt_2$. Similarly, since $p_{N+1}\eqgran q_{N+1}$, we have $\stmt_1\eqgran\stmt_1'\OR q_{N+1}$. We conclude that $\stmt_1\eqgran\stmt_2$. This completes the proof that $2\Rightarrow3$.

The claim $3\Rightarrow1$ is immediate since 1 is a special case of 3. 
\end{proof}

\begin{remark}
The above observations and propositions show that MCP and comparability of possibilities imply equigranular possibilities, but equigrangular possibilities and totality do \emph{not} imply MCP.

Further, MCP alone implies neither totality nor equigranular possibilities. Indeed, consider a theoretical domain with only two possibilities and a fineness ordering which has two unequal successors to the fineness of $\impossibility$. This satisfies the MCP but fails the other conditions. 
\end{remark}

\begin{prop}\label{uniformconsequence}
    Let $\tdomain$ be uniform. Let $\stmt, \stmt_1 \in \tdomain$ such that $\stmt \sfiner \stmt_1$. Then there exist $\stmt_2 \in \tdomain$ such that $\stmt_2 \narrower \stmt_1$ and $\stmt \eqgran \stmt_2$.
\end{prop}

The start of a proof....
Let $S = \{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$. We will find a maximal element with respect to fineness (OR narrowness??).

The difficulty here is that for Zorn we need to find an upper bound of an arbitrary chain which also lies in $S$. But such an upper bound is entirely unclear on how to obtain without using the proposition to begin with. Uniformity is difficult to make much use of for infinite domains (beyond that $S$ is nonempty).

This set $S$ is nonempty, since any possibility narrower than $\stmt_1$ must also be finer than $\stmt$ by uniformity, and the statement $\stmt_1$ must not be the impossibility since it is strictly coarser than the statement $\stmt$.

Now, given any chain (with respect to narrowness), consider its disjunction. But again this may be coarser than $\stmt$!

Note: problem with chains for fineness. Taking the disjunction of a chain with respect to fineness can yield something much coarser (i.e. outside of $\{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$) if they do not overlap.

Idea: given chains in fineness, can we construct new chains which are, at each level, equigranular, but now they are also chains in narrowness?

Lemma: If $C$ is a chain in fineness and $\stmt$ is such that $\stmt[t]\finer\stmt$ for all $\stmt[t]\in C$, then there exists a set $N$ and a bijection $f:C\to N$ such that $N$ is a chain in narrowness, $\OR N \narrower \stmt$, and for all $\stmt[t]\in C$, $\stmt[t]\eqgran f(\stmt[t])$. 
Use this to move chains in fineness over to chains in narrowness and then can find maximal elements in the set $S$ above ``up to equigranular equivalence classes".

Problem: this is actually much harder. Prop 6 is the case of this Lemma for a chain of length 2 and the top of the chain being equal to $\stmt$.

\section{Notes}

One condition for the measure to be unique given a unit statements is that all points (i.e. possibilities in our framework) are equigranular. In fact, consider a domain with just three points not equigranular to each other, but each finer than the other. We can construct any measure that satisfies the requirement by picking three numbers and assigning them accordingly, since we can't determine ``how bigger'' is one point with respect to the other.

\subsection{Finding a probability measure}

Our first goal is to find a measure which is compatible with the fineness preorder. The idea is that fineness tells us which sets are ``larger" (less precise) than each other, regardless of any overlap.  A measure will quantify this and allow us to assign numbers to the sizes.

The paper \cite{probexistence} seems to be extremely relevant here. The basic idea in this paper is to find necessary and sufficient conditions for the existence of a probability measure on a Boolean algebra (e.g. of subsets of some set) which is compatible with a total pre-order. There are a few points to keep in mind.

First, we need a \emph{total} pre-order. This means for every pair of elements, one is finer than the other. This is certainly not the case for fineness in general: color and position, for example, cannot be compared. This is the goal of defining the \emph{unit} above. The unit serves as a baseline statement against which others can be compared for fineness. For example, a statement about position singles out length as the dimension to compare, or a statement about numbers of objects specifies ``count" as the dimension to compare (e.g. the discrete unit of number of objects).

Next, we need the three properties specified in the paper, to be written up later! We are optimistic that these three properties are present in our situation.

There is one additional caveat. The measure constructed by the main theorem in the above-mentioned paper is always a probability measure. This means the whole space has finite measure. For example, one could put the Lebesgue measure on $(0,1)$ but under a homeomorphism $(0,1)\to\mathbb{R}$ induce a finite (probability) measure on  $\mathbb{R}$. If we chose the full set as a unit, we are exactly looking for a probability measure. If we pick a unit that is not the full set, we can also "constrain" the problem to that subset, so that we still have a probability measure. Then we may be able to "stich together" the measures constructed on all equigranular sets assuming they cover the full space.

\subsection{Applying results of the paper \cite{villegas}}

In this part, we will assume that $\tdomain_{\stmt[u]}$ is a subset of the theoretical domain, $\stmt[u]\in \tdomain_{\stmt[u]}$ is a distinguished element (we'll refer to it as the \emph{unit} when needed), and there is a total pre-order $\leq_{\stmt[u]}$ defined on $\tdomain_{\stmt[u]}$. We will write $=_{\stmt[u]}$ for $\leq_{\stmt[u]}$ and $\geq_{\stmt[u]}$, and we write $<_{\stmt[u]}$ for $\leq_{\stmt[u]}$ and not $=_{\stmt[u]}$. We will use the word ``equiprobable" for $=_{\stmt[u]}$ since it means something different from equigranular (since $\leq_{\stmt[u]}$ is a different ordering from fineness, even though it will be derived from fineness). We assume $\tdomain_{\stmt[u]}$ is closed under complement and countable union (and hence countable intersection), and that there are two additional elements, $\certainty$ and $\impossibility$, which are maximal and minimal in this ordering, respectively.

\begin{remark}
In defining the subset $\tdomain_{\stmt[u]}$ and the total preorder $\leq_{\stmt[u]}$, we will have to be very careful. We want to ``squash" together all the sets which should be measure zero in this unit (e.g. anything ``lower-dimensional"), ``blow up" all the ``infinite measure" subsets (e.g. ``higher dimensional"), and ensure that everything is compatible under this order, so there can be only one ``type" of property under consideration. 
\end{remark}

We write out the main axioms and statements from \cite{villegas} which interest us in this situation, in the notation established earlier. On page 1789, we have the following:

Axiom of Pre-order:
\begin{itemize}
    \item Totality: If $\stmt_1,\stmt_2\in \tdomain_{\stmt[u]}$, then $\stmt_1 \leq_{\stmt[u]} \stmt_2$ or $\stmt_2\leq_{\stmt[u]} \stmt_1$
    \item Reflexivity: For all statements $\stmt\in \tdomain_{\stmt[u]}$, we have $\stmt\leq_{\stmt[u]}\stmt$
    \item Transitivity: For all statements $\stmt_1,\stmt_2,\stmt_3\in \tdomain_{\stmt[u]}$, if $\stmt_1\leq_{\stmt[u]}\stmt_2$ and $\stmt_2\leq_{\stmt[u]}\stmt_3$, then $\stmt_1\leq_{\stmt[u]}\stmt_3$.
    \item Existence of max and min: We have $\impossibility\leq_{\stmt[u]}\certainty$ and \emph{not} $\certainty\leq_{\stmt[u]}\impossibility$. For all $\stmt\in \tdomain_{\stmt[u]}$ we have $\impossibility\leq_{\stmt[u]}\stmt\leq_{\stmt[u]}\certainty$
\end{itemize}

The above list is simply the requirement that $\leq_{\stmt[u]}$ is a total pre-order and that there are a maximal and minimal element.

Axiom of Monotony: 
Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$. Assume that $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, and $\stmt_2\ncomp \stmt_4$ (i.e. $\stmt_2\AND \stmt_4 =_{\stmt[u]} \impossibility$). Then we have $\stmt_1\OR \stmt_3 \leq_{\stmt[u]} \stmt_2\OR \stmt_4.$

Monotony guarantees that given two incompatible statements, each ``bigger" than another statement, then their disjunction is still bigger than the disjunction of the smaller statements.

The following are consequences of the two axioms above. Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
\begin{itemize}
    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ and $\stmt_2<_{\stmt[u]} \stmt_3$, then $\stmt_1<_{\stmt[u]}\stmt_3$. 
    \item If $\stmt_1\narrower \stmt_2$ then $\stmt_1\leq_{\stmt[u]} \stmt_2$
    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, $\stmt_2\leq_{\stmt[u]} \stmt_4$, and $v\leq_{\stmt[u]} \stmt_1$, then $\stmt_2 \AND \NOT \stmt_1 \leq_{\stmt[u]} \stmt_4 \AND \NOT \stmt_3$
    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ then $\NOT \stmt_2 \leq_{\stmt[u]}\NOT \stmt_1$
\end{itemize}

Next, we will need a more subtle axiom in order to make sense of infinite combinations of statements.

\begin{remark}
The \emph{fineness} relation $\finer$ should satisfy all of the axioms of pre-order and the axiom of monotony, except for totality. We restrict to $\tdomain_{\stmt[u]}$ in order to get totality. 
\end{remark}

Axiom of Monotone Continuity:
Let $\{\stmt_i : i\in \mathbb{Z}^+\}$ be a sequence of statements which is monotone increasing with respect to narrowness (containment), and let $\stmt = \cup_i\stmt_i$ be the disjunction. Let $t\in \tdomain_{\stmt[u]}$ such that $\stmt_i\leq_{\stmt[u]} t$ for all $i$. Then $\stmt\leq_{\stmt[u]} t$.

In other words, an infinite disjunction can only strictly exceed a statement in coarseness if the disjunction of a finite subset of them already exceeds it. Here is an interesting consequence of the combination of the above axioms:

\begin{prop}[Corollary to Lemma 1 in \cite{villegas}]
\label{infinitenull}
Let $\{\stmt_i : i\in \mathbb{Z}^+\}$ be an infinite collection such that $\stmt_i\AND\stmt_j =_{\stmt[u]} \impossibility$ for all $i\neq j$, and $\stmt_i =_{\stmt[u]} \stmt_j$ for all $i,j$. Then it must be the case that $\stmt_i=_{\stmt[u]}\impossibility$ for all $i$. 
\end{prop}
This says that infinite collections of equiprobable statements with null overlaps must be null themselves.

An \emph{atom} is a statement $\stmt\in S$ such that $\impossibility<_{\stmt[u]}\stmt$ and there does not exist any statement $\stmt_2\in S$ such that $\impossibility<_{\stmt[u]} \stmt_2<_{\stmt[u]} \stmt$. In particular, Theorem 5 in \cite{villegas} \S3 includes the fact that $\tdomain_{\stmt[u]}$ is atomless if and only if every statement can be partitioned into two equiprobable statements. Theorem 3 in \S4 states that if $\tdomain_{\stmt[u]}$ is atomless, then there exists a unique probability measure compatible with $\leq_{\stmt[u]}$, and it is countably additive!

\subsection{Using Monotone Continuity}

The core of the proof of Theorem 3 in Section 4 of \cite{villegas} comes from Theorem 5 (iv) in Section 3. In the terminology of Villegas, a \emph{qualitative probability $\sigma$-algebra} is a $\sigma$-algebra of subsets of a space with an ordering that satisfies the Axiom of Pre-order, the Axiom of Monotony, and the Axiom of Monotone Continuity. We also need the language of \emph{atoms}, defined earlier. We will reproduce some of the ideas in the proofs of these theorems in the language of statements and our theoretical domain, while attempting to keep track of the usages of Monotone Continuity. The hope is that there is a weakened version of the axiom that gets us the desired result, or that we can build the ordering $\leq_{\stmt[u]}$ with Monotone Continuity but it can still distinguish the infinite sets described earlier.

Let $A$ be a statement in $\tdomain_{\stmt[u]}$ with $A>_{\stmt[u]} \impossibility$. A \emph{minor incomplete partition} of $A$ is a collection $M_A$ of statements with the following properties:
\begin{itemize}
    \item We have $\stmt>_{\stmt[u]}\impossibility$ and $\stmt \narrower A$ for all $\stmt\in M_A$, that is, they are non-null and narrower than $A$
    \item For all $\stmt_1,\stmt_2\in M_A$, we have $\stmt_1\AND\stmt_2 =_{\stmt[u]} \impossibility$, that is, they overlap only on statements ``squashed" to zero. 
    \item The disjunction of $M_A$ is finer than its complement in $A$: $$\bigOR_{\stmt\in M_A}\stmt \finer (A\AND \NOT\bigOR_{\stmt\in M_A}\stmt).$$ This is the condition of being \emph{minor}.
    %UM_A finer than A - UM_A^C
\end{itemize}

It is shown in \cite{villegas} that any (minor) incomplete partition is at most countable (Theorem 3). Rather than give the full proof of this fact, we note that it follows from a large collection of lemmas, the most interesting of which is the following, which we will prove below:

\begin{prop}[Lemma 1 from \cite{villegas}]
If $\{\stmt[a]_i: i\in I\}$ is an infinite set of statements such that (1) $\stmt[a]_i \AND \stmt[a]_j =_{\stmt[u]}\impossibility$ for all $i\neq j$ and (2) $\stmt[a]_i>_{\stmt[u]}\impossibility$ for all $i$, then the following holds. Given a statement $\stmt>_{\stmt[u]}\impossibility$, only finitely many of the $\stmt[a]_i$'s have $\stmt[a]_i>_{\stmt[u]}\stmt$.  
\end{prop}
\begin{proof}
By way of contradiction, let $\stmt[a]_1,\ldots$ be a sequence of statements violating the conclusion (i.e. infinitely many statements $\stmt[a]_i$ with $\stmt[a]_i>_{\stmt[u]}\stmt>_{\stmt[u]}\impossibility$. Consider the new sequence $$\stmt_i = \bigOR\limits_{n=i}^{\infty}\stmt[a]_n.$$ This sequence is monotonely decreasing to the impossibility in narrowness (and hence in fineness and in $>_{\stmt[u]}$), but we also have $\stmt_i\geq_{\stmt[u]}\stmt$ for all $i$. By monotone continuity (the contrapositive), we have $\stmt[s]\leq_{\stmt[u]}\impossibility$, a contradiction. 
\end{proof}

A quick corollary of this is Proposition \ref{infinitenull}, above. Using minor incomplete partitions, we prove the following.

\begin{prop}[Theorem 4, \S3 \cite{villegas}]
\label{split}
If there are no atoms, then for every statement $\stmt$ there exist statements $\stmt_1$ and $\stmt_2$ such that: $\stmt_1\ncomp \stmt_2$, $\stmt_1\equ\stmt_2$, and $\stmt_1\OR\stmt_2 = \stmt$. In other words, $\stmt$ can be partitioned into two equiprobable (note: I remember we don't like this word but I don't remember if we replaced it or not) statements. 
\end{prop}
\begin{proof}
Let $S$ be the set of minor incomplete partitions of $\stmt$, and endow it with a partial order by inclusion. Given a totally ordered chain in $S$, its union is still a minor incomplete partition, and is thus an upper bound for the chain. We may apply Zorn's lemma to find a maximal element of $S$, the disjunction of which we call $\stmt_1$. We claim that $\stmt_1$ and $\stmt_2:= \stmt\AND\NOT\stmt_1$ are equiprobable, incompatible statements. Incompatibility is immediate from the definition. For equiprobability, first note that by definition of being a minor incomplete partition we have $\stmt_1\lequ\stmt_2$. Next, if $\stmt[t]_1,\ldots$ is a monotone decreasing (with respect to $\lequ$) sequence of statements $\stmt[t]_n>_{\stmt[u]}\impossibility$ narrower than $\stmt_2$ and converging to a zero-comparable event, we have that $\stmt_1\OR\stmt[t]_n>_{\stmt[u]}\stmt_2\OR\stmt[t]_n$ since $\stmt_1$ is maximal. We needed no atoms in order to guarantee this sequence exists (Lemma 3 in Section 3 of \cite{villegas}). It follows by monotone continuity that $\stmt_1\geq_{\stmt[u]} \stmt_2$. 
\end{proof}

We will also need a few types of \emph{random variables} as defined in \cite{villegas} \S3. Let $A({\stmt[u]}) = \{x \in X : x \comp \stmt[u]\}$ be the set of possibilities associated with $\stmt[u]$. This is a subset of the collection of possibilities which will serve as the ``sample space" after picking a unit $\stmt[u]$.

A \emph{uniform sequence of finitely-valued random variables} is a sequence of functions (random variables in the traditional sense) $X_1,X_2,\ldots: A({\stmt[u]})\to E$ where $E$ is a finite set, with the following property: for all $\{x_1,\ldots,x_n\}\subseteq E$, all the events $$\{p\in A({\stmt[u]}): X_i(p) = x_i, i=1,\ldots,n\}$$
are equiprobable ($\equ$ with each other).

Notice that monotone continuity allows us to let $n\to\infty$ to define events where the whole sequence is specified with the resulting events all being equiprobable. By Proposition \ref{infinitenull} above, all such events are null.

Building on these sequences of random variables, we will use \emph{uniform random variables} to define probability measures. Let $X:A({\stmt[u]})\to[0,1]$ be a random variable, and let $I$ and $J$ be intervals contained in $[0,1]$. We say $X$ is a uniform random variable if $X^{-1}(I)\lequ X^{-1}(J)$ if and only if $|I|\leq|J|$ for all such $I,J$.

From Proposition \ref{infinitenull}, $X^{-1}(a)$ is a zero-comparable event for all $a\in[0,1]$, and if $0=a_0<a_1<\cdots<a_n=1$, then $X^{-1}([a_i,a_{i+1}))$ are a finite partition of $\tdomain$.

Next, we will pull many of the above concepts together to show how we obtain a probability measure on $A({\stmt[u]})$ realized as a function on $\tdomain$ compatible with $\lequ$.

\begin{prop}\label{uniformexist}
If $\tdomain_{\stmt[u]}$ has no atoms with respect to $\lequ$, then there exists a uniform random variable.
\end{prop}
\begin{proof}
By Proposition \ref{split} we may split any statement into two equiprobable, null-compatible sets. Starting with the statement $\stmt[u]$ (here, the ``universe" which we want to have probability 1), repeatedly divide into two statements in stages. At stage $n$ (starting with $n=0$), we will have statements $\stmt[u]^{(n)}_1,\ldots,\stmt[u]^{(n)}_{2^n}$. Now, take the disjunction of alternating statements in each collection so that there is a ``right" and ``left" (odd and even index) half at each stage. For example, at stage $n=1$, the odd and even halves are simply $\stmt[u]^{(1)}_1$ and $\stmt[u]^{(1)}_2$; at stage $n=2$ we have $\stmt[u]^{(2)}_1\OR\stmt[u]^{(2)}_3$ and $\stmt[u]^{(2)}_2\OR\stmt[u]^{(2)}_4$.

Define a sequence of random variables $X_n:A(\stmt[u])\to\{0,1\}$ by $X_n(a) = 1$ if and only if $a$ is in the odd half in stage $n$. This gives each $a\in A(\stmt[u])$ a binary ``address". By the construction, it follows that this is a uniform sequence of finitely-valued random variables (defined above).

Now, we define a uniform random variable $X:A(\stmt[u])\to[0,1]$. For $a\in A(\stmt[u])$, define $X(a) = \sum_{n=1}^{\infty}2^{-n}X_n(a)$. In other words, $X(a)$ is the binary expansion $0.X_1(a)X_2(a)\ldots_2$.

One can show immediately that $X^{-1}(I)\lequ X^{-1}(J)$ if and only if $|I|\leq |J|$ when $I$ and $J$ have endpoints on numbers with finite binary expansions (since $X_1,\ldots,X_n$ are uniform), and by monotone continuity, the uniformity condition extends to arbitrary intervals in $[0,1]$. 
\end{proof}

Finally, we reach our goal of building a probability measure on $A(\stmt[u])$.

\begin{prop}
If $\tdomain_{\stmt[u]}$ has no atoms with respect to $\lequ$, then there exists a unique compatible probability measure, and it is countably additive. 
\end{prop}
\begin{proof}
By Propsition \ref{uniformexist}, let $X:POSS_{\stmt[u]}\to[0,1]$ be a uniform random variable.
%Define $F:[0,1]\to\tdomain$ by $F(x) = X^{-1}([0,x])$. It is straightforward to see that $F$ is monotone in $x$ (with respect to $\lequ$), and that if $x_n\to x$, then $\lim_nF(x_n) \equ F(x)$. %Started writing this part from paper, but it is not directly used in construction. Will add details later as needed. 
We claim that for any statement $\stmt\in\tdomain_{\stmt[u]}$, there exists a unique number $p\in[0,1]$ such that $\stmt\equ X^{-1}([0,p])$. Define $P:\tdomain_{\stmt[u]}\to[0,1]$ by $P(\stmt)= p$. This defines a probability measure compatible with $\lequ$. Details forthcoming as needed....
\end{proof}

Ideas involved: Defines a function $X:A(\stmt[u])\to [0,1]$ in a way that is consistent with $\lequ$ by having ``larger" sets of possibilities (i.e. coarser statements) corresponding to larger intervals in $[0,1]$. Then, the measure of any other statement is given by finding an equiprobable statement which is the preimage of $[0,p]$ under $X$ and defining the measure to be $p$.

Wish: find a way to prove the above statements in an order-theoretic way. In particular, we wish to show that the order type of the equigranular-equivalence classes is equivalent to (a subset of) $\mathbb{R}$. Sufficient conditions to be isomorphic to $\mathbb{R}$ are (1) lacking endpoints (2) complete (i.e. has Least Upper Bound property) (3) has dense countable subset. The use of random variables may be very similar to an approach using order theory. For example, the transition from a uniform sequence of finitely-valued random variables into a (single) uniform random variable makes use of a countable dense subset.

\subsection{Example with lack of monotonicity}

Consider a Euclidian two dimensional space charted by $\{x,y\}$. Let $U_{i,j}$ be the square in the interval $x \in (i, i+1)$ and $y \in (j, j+1)$. Let $V$ be the vertical (infinite) strip in the interval $x \in (0, 1)$ and $W$ the vertical strip in the interval $x \in (0, 2)$. Define the sequence $W_i = \{ U_{0,0}, U_{1,0}, U_{0,1}, U_{1,1}, U_{0,2}, U_{1,2}, U_{0,3}, U_{1,3}, ... \}$. We have $\bigcup\limits_0^\infty W_i = W$. This means $\bigcup\limits_0^n W\finer V$ for all finite $n$ but $\bigcup\limits_0^\infty W_i = W \coarser V$.

\subsection{Some thoughts on Example in 3.3}

The construction above hints at some fundamental concepts in measure theory which will be important for us to consider.

Notice that the two sets $V$ and $W$ are both of infinite Lebesgue measure (area). With respect to Lebesgue measure, monotone continuity is immediate since if $\bigcup\limits_0^n W\finer V$ for all $n$ then $\lim\limits_{n\to\infty}\mu(\bigcup\limits_0^n W) = \mu(W)\finer V$. Even the above example does not violate this, since there, both would be ``equivalent" in fineness, namely they are infinitely coarse.

Our goal is to have a measure which is able to distinguish sets which may be of infinite Lebesgue measure as well. This will be very difficult to fully capture. Consider the following procedure. Starting with the vertical strip $V$, slide all the squares at even $y$-values to the right by one, and then allow the squares to collapse down to rest on the square below. Now, by simply rearranging regions, we have (almost) a cover of the wider strip $W$. The only missing points are the line $x=1, y>0$. If instead we used $(0,1]$ and $(0,2]$ for the strips, this procedure would cover them exactly. One could continue and cover the entire plane with similar translations.

Here is one approach to distinguish $V$ and $W$, and similar methods will enable one to distinguish many more infinite-measure subsets. Let $S_n$ be the $2n\times 2n$ square centered the origin with vertices at $(\pm n, \pm n)$. Then for all $n$, $V\cap S_n$ has area $n$, and $W\cap S_n$ has area $2n$. This holds for all $n$ and so we might use something like this to think about how we would distinguish infinite areas. Notice that this is in some sense writing $V$ and $W$ as countable unions of nested sets in a standardized way (e.g. $V = \bigcup\limits_n V\cap S_n$) and then comparing the finite-area subsets. One could choose shapes other than $S_n$ for this purpose. Another way to distinguish them would be to look at the limit $\lim\limits_{n\to\infty}\text{Area}(V\cap S_n)/\text{Area}(S_n)$; in this case they are both 0 but for a different choice of $S_n$ one could find an asymptotic difference. 
The point of the above discussion is to try to gain a better idea of which (infinite/unbounded) sets we want to be related by $\finer$. It is not clear even for ``nice" regions of $\mathbb{R}^2$, let alone at the full generality we are seeking.

Regardless, the goal is not just to distinguish infinite-Lebesgue-measure sets. We need to understand when monotone continuity is relevant and what happens without it.

\subsection{Infinitesimals}

\subsubsection{Notes on deFinetti and Archimedean property }

On Page 21 de Finetti specifically says that $P(E) \geq P(E’)$ implies $E \geq E’$ but not vice-versa. And that we can have $P(A)=P(B)=0$ even if $A<B$ if $A$ is impossible and $B$ is not. So, the ordering he has in mind is fineness for us. And then he says that this means that the 4 postulates give a non-Archimedean structure. But you can make it Archimedean by ignoring the probabilities that are “infinitesimally small”: those that multiplied by a number, however big, will never become the certainty. This whole thing is spot on.

From a cursory look at the Archimedean property (\url{https://en.wikipedia.org/wiki/Archimedean_property})
this is typically discussed to define notions of infinitesimals in an ordered group. Most notably, the real number (and any complete ordering) will satisfy the Archimedean property.

\subsubsection{Infinitesimals}

Having seen that, we can define infinitesimal in the following way.

\begin{defn}\label{infinitesimals}
Let $\stmt[u],\stmt[s] \in \tdomain$. Then $\stmt$ is \textbf{infinitesimal} with respect to $\stmt[u]$ if for any natural number $n$ we can find a set of $n$ statements $\{ \stmt_i \}_{i=1}^n$ such that:
\begin{enumerate}
    \item $\stmt \finer \stmt_i$ for all $i$
    \item $\stmt_i \ncomp \stmt_j$ for all $i \neq j$
    \item $\bigOR\limits_{i=1}^n \stmt_i \finer \stmt[u]$ 
\end{enumerate}
In words, $\stmt$ is infinitesimal with respect to $\stmt[u]$ if for any number, there is a disjunction of that many broader mutually incompatible statements $\stmt_i$ is narrower than $\stmt[u]$.
\end{defn}

\begin{coro}
    The impossibility $\impossibility$ is an infinitesimal with respect to any $\stmt[u] \in \tdomain$.
\end{coro}

\begin{proof}
    Let $\stmt[u] \in \tdomain$ and consider $\{ \stmt_i \}_{i=1}^n$ where $\stmt_i \equiv \impossibility$ for all $i$. Then $\impossibility \finer \impossibility \equiv \stmt_i$ for all $i$. Since $\impossibility \ncomp \impossibility$, $\stmt_i \ncomp \stmt_j$ for all $i$ and $j$. Lastly, $\bigOR\limits_{i=1}^n \stmt_i \equiv \impossibility \finer \stmt[u]$. This means that $\{ \stmt_i \}_{i=1}^n$ realizes infinitesimality for $\impossibility$, which is an infinitesimal of $\stmt[u]$.
\end{proof}

\begin{prop}\label{infinitesimal_conjunction}
Let $\stmt[v] \in \tdomain$ be an infinitesimal with respect to $\stmt[u] \in \tdomain$. Then any $\stmt \finer \stmt[u]$ is also an infinitesimal with respect to $\stmt[u]$.
\end{prop}
\begin{proof}
 Since $\stmt[v]$ is an infinitesimal with respect to $\stmt[u]$, for any $n$ we can find a set of statements $\{ \stmt_i \}_{i=1}^n$ that realizes its infinitesimality. We have $\stmt \finer \stmt[v] \finer \stmt_i$ for all $i$. The other two properties do not depend on $\stmt[v]$. Therefore $\{ \stmt_i \}_{i=1}^n$ realizes infinitesimality for $\stmt$ as well so $\stmt$ is an infinitesimal with respect to $\stmt[u]$.
\end{proof}

\begin{coro}
The conjunction of two infinitesimals with respect to $\stmt[u]$ is infinitesimal with respect to $\stmt[u]$.
\end{coro}
\begin{proof}
Let $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ be infinitesimal with respect to $\stmt[u]$. Since $\stmt[v]_1 \AND \stmt[v]_2 \narrower \stmt[v]_1$, $\stmt[v]_1 \AND \stmt[v]_2 \finer \stmt[v]_1$ and therefore $\stmt[v]_1 \AND \stmt[v]_2$ by \ref{infinitesimal_conjunction}.
\end{proof}

\begin{prop}
The disjunction of two infinitesimals with respect to $\stmt[u]$ that are comparable\footnote{Is comparability required? Can it be proved one way or another?} to each other is infinitesimal with respect to $\stmt[u]$.
\end{prop}
\begin{proof}
Let $\stmt[v],\stmt[w]$ be infinitesimal with respect to $\stmt[u]$. Without loss of generality, suppose $\stmt[v] \finer \stmt[w]$. Let $n>0$. Let $\{\stmt[w]_i\}_{i=1}^{2n}$ be a collection realizing infinitesimality for $2n$ for $\stmt[w]$. Define $\stmt_j = \stmt[w]_j\OR\stmt[w]_{j+n}$ for $j=1,\ldots,n$. Notice that the $\stmt_j$'s are mutually incompatible, and their disjunction must be finer than $\stmt[u]$. Further, since $\stmt[v]\finer\stmt[w]$ we have that $\stmt[v]\finer\stmt[w]_i$ for all $i$. It follows that:
$$
\stmt[v]\OR\stmt[w] \finer \stmt[w]_j\OR\stmt[w]_{j+n} = \stmt_j
$$
for all $j=1,\ldots,n$. Thus the collection $\{\stmt_j\}_{j=1}^n$ realizes $\stmt[v]\OR\stmt[w]$ being infinitesimal with respect to $\stmt[u]$ for $n$. We conclude that $\stmt[v]\OR\stmt[w]$ is infinitesimal with respect to $\stmt[u]$. 
\end{proof}

\begin{prop}[Infinitesimality is consistent with fineness]
Let $\stmt[u],\stmt[v]$ be statements such that $\stmt[v]\finer\stmt[u]$. Let $\stmt$ be a statement $\stmt\narrower\stmt[u]\AND\stmt[v]$ such that $\stmt$ is infinitesimal with respect to $\stmt[v]$. Then $\stmt$ is infinitesimal with respect to $\stmt[u]$. 
\end{prop}
\begin{proof}
Let $n>0$ be a natural number, and let $\{ \stmt_i \}_{i=1}^n$ be as in the definition of infinitesimal for $\stmt$ with respect to $\stmt[v]$. Then $\{ \stmt_i \}_{i=1}^n$ also satisfies the requirements for infinitesimal with respect to $\stmt[u]$. The first two requirements are unchanged, and the third is easy to see:
$$
\bigOR\limits_{i=1}^n \stmt_i \finer\stmt[v]\finer \stmt[u]
$$
\end{proof}

As an immediate corollary to the above, we have:
\begin{prop}
Infinitesimality is consistent with $\narrower$.
\end{prop}

We wish to also show some form of compatibility in the reverse direction. That is, if we shrink by a ``finite" amount, the sets which were previously infinitesimal should remain so (of course, if we shrink by an ``infinite" amount, e.g. by dropping a dimension, previously infinitesimal sets may no longer be so).

\begin{prop}
Let $\stmt[u],\stmt[v]$ be statements such that $\stmt[v]$ is not infinitesimal with respect to $\stmt[u]$. Let $\stmt$ be a statement $\stmt\narrower\stmt[v]$ such that $\stmt$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt$ is infinitesimal with respect to $\stmt[v]$. 
\end{prop}
\begin{proof}
Let $\stmt[u]$, $\stmt[v]$, and $\stmt$ be as in the Proposition statement. We will show that $\stmt$ is infinitesimal with respect to $\stmt[v]$. To do so, we will show that for any $n>0$, we can find a set of $n$ statements $\{\stmt_i\}_{i=1}^n$ such that: 
\begin{enumerate}
    \item $\stmt \finer \stmt_i$ for all $i$
    \item $\stmt_i \ncomp \stmt_j$ for all $i \neq j$
    \item $\bigOR\limits_{i=1}^n \stmt_i \finer \stmt[v]$ 
\end{enumerate}
We will show this for arbitrary $n$, which will be sufficient to prove that $\stmt$ is infinitesimal with respect to $\stmt[v]$.

First, since $\stmt[v]$ is not infinitesimal with respect to $\stmt[u]$, there exists some $N>0$ such that for any set of $N$ statements, at least one of the conditions of infinitesimality of $\stmt[v]$ with respect to $\stmt[u]$ fails.

Let $n>0$. Let $\{\stmt_i\}_{i=1}^{nN}$ be a set of $n\cdot N$ statements realizing infinitesimality of $\stmt$ with respect to $\stmt[u]$ (i.e. they are mutually incompatible, broader than $\stmt$, and their disjunction is narrower than $\stmt[u]$).

{\bf By totality of $\finer$}\footnote{What we needed here is that there is a partition of the collection into $N$ sets of size $n$ such that for one of them, their disjunction is finer than each of $N-1$ disjunctions of the other subsets in the partition.}, we may assume the collection $\{\stmt_i\}_{i=1}^{nN}$ is ordered by fineness ($\stmt_1\finer\stmt_2\finer\cdots$). Consider the finest $n$ statements, that is, consider the collection $\{\stmt_i\}_{i=1}^{n}$. We will show that this collection satisfies the conditions for $\stmt$ being infinitesimal with respect to $\stmt[v]$ for $n$ statements.

By construction, this satisfies properties 1 and 2 for infinitesimality of $\stmt$ in $\stmt[v]$, since it is a subset of a collection which already satisfied those properties.

Now suppose it fails property 3\footnote{We will use totality since we need $\NOT \big(\stmt[x]\sfiner\stmt[y]\big) \implies \stmt[x]\coarser\stmt[y]$.}, that is, suppose 
$$
\bigOR\limits_{i=1}^n \stmt_i \scoarser\stmt[v].
$$
We will show that this will lead to a contradiction by implying that there is a collection of statements of size $N$ realizing $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$.

Consider now the collection of $N$ statements $\{\stmt[a]_j\}_{j=1}^N$ defined by $$\stmt[a]_j = \bigOR\limits_{i=n(j-1)+1}^{nj}\stmt_i.$$
These statements are built from the collection $\{\stmt_i\}_{i=1}^{n}$ by taking $N$ disjunctions each consisting of $n$ of the statements. Notice that $\stmt[a]_1$ is the disjunction of the collection $\{\stmt_i\}_{i=1}^{n}$ considered above.

First, this collection is mutually incompatible, since it is made up of disjunctions of mutually incompatible statements. This is property 2 for $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$.

Next, these must satisfy property 1 for $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$. To see why, notice that by our assumption, we have $\stmt[a]_1\scoarser\stmt[v]$, and by construction, we have $\stmt[a]_j\coarser\stmt[a]_1$ for all $j=1,\ldots,N$, and so $\stmt[v]\finer\stmt[a]_j$ for all $j=1,\ldots,N$.

Finally, they also must satisfy property 3: $$\stmt[a]_1\OR\cdots\OR\stmt[a]_N = \bigOR\limits_{i=1}^{nN}\stmt_i\finer\stmt[u].$$

Thus, this new collection violates the assumption that no set of $N$ statements satisfies all three properties for $\stmt[v]$ being infinitesimal in $\stmt[u]$.

{\bf By totality of $\finer$},\footnote{Here we are using totality as specified in the previous footnote.} we conclude that we must have $\bigOR\limits_{i=1}^n \stmt_i \finer\stmt[v]$ as required, and so the collection $\{\stmt_i\}_{i=1}^{n}$ realizes $\stmt$ being infinitesimal with respect to $\stmt[v]$ for $n$ statements. Since $n$ was arbitrary, it must be the case that $\stmt$ is infinitesimal with respect to $\stmt[v]$. 
\end{proof}

\subsubsection{Equivalence at u scale}

\def\equivu{\cong_{\stmt[u]}}

\begin{defn}
Let $\stmt_1, \stmt_2 \in \tdomain$. We say $\stmt_1$ is \textbf{equivalent to $\stmt_2$ at $\stmt[u]$ scale}, noted $\stmt_1 \equivu \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$.
\end{defn}

\begin{prop}
Equivalence at $\stmt[u]$ scale is an equivalence class.
\end{prop}
\begin{proof}
For reflexivity, let $\stmt \in \tdomain$. We have $\stmt \OR \impossibility \equiv \stmt \OR \impossibility$. Since $\impossibility$ is an infinitesimal for all $\stmt[u]$ (TODO: add proposition), $\stmt \equivu \stmt$. For symmetry, note that the definition itself is symmetrical. For transitivity, let $\stmt_1 \equivu \stmt_2 \equivu \stmt_3$. Then we have $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$ and $\stmt_2 \OR \stmt[w]_2 \equiv \stmt_3 \OR \stmt[w]_3$.
\end{proof}

\subsubsection{Greatness in $\stmt[u]$}

First define ``as big in $\stmt[u]$'' $\equ$.

\begin{defn}
Let $\stmt_1, \stmt_2 \in \tdomain$. Then $\stmt_1$ is \textbf{as big as} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \equ \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$.
\end{defn}

\begin{prop}
Suppose $\tdomain$ is uniform. Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$, then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
\end{prop}

We start with a simpler case:

\begin{prop}
Let $\stmt_1,\stmt,\stmt[v]_1\in\tdomain$ such that $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$, and suppose $\stmt[v]_1$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt\equ\stmt_1$. 
\end{prop}
\begin{proof}[Using Prop \ref{uniformconsequence}]
If both of the $\finer$'s are actually equigranular, there is nothing to show. Next consider the case that $\stmt_1\sfiner\stmt\finer\stmt_1\OR\stmt[v]_1$. Then by Proposition \ref{uniformconsequence} there exists $\stmt_2$ such that $\stmt_2\narrower\stmt$ and $\stmt_2\eqgran\stmt_1$. But how do we put this together?? What about the even simpler version of the lemma with $\narrower$ instead of $\finer$? Just use $\stmt\OR (\stmt_1\OR\stmt[v]_1 \AND \NOT\stmt)$.
\end{proof}

This seems problematic already:

We have $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$. We need to find some $\stmt[w],\stmt[w]_1$ infinitesimal with respect to $\stmt[u]$, such that $\stmt\OR\stmt[w]\eqgran\stmt_1\OR\stmt[w]_1$. Since we don't know anything about the overlap of $\stmt$ and $\stmt_1$, we can't use any operations like $\stmt\AND\NOT\stmt_1$ to find something small. We need to know that additional statements exist in order to find $\stmt[w]$ and $\stmt[w]_1$.

For example, what if every infinitesimal statement was narrower $(\narrower)$ than $\stmt$? In this case, we could have something like $\stmt_1\sfiner\stmt\sfiner\stmt_1\OR\stmt[x]$ for all statements (including all infinitesimals) $\stmt[x]$. It would then not be the case that $\stmt\equ\stmt_1$.

\begin{prop}[Need something like this but not clear if correct!]
Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$. Then there exist statements  $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$, infinitesimal in $\stmt[u]$, such that:
$$
\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
$$
\end{prop}

Note that the existence of the right combination of incompatible statements is sufficient. However, that is stronger than we need and may be harder to prove. Problem: we have very few propositions about existence of statements! It is hard to ensure certain ones exist.

Consider the following (extreme) case. Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$, but the infinitesimals which realize these relationships are $\stmt[v]_1$ and $\stmt[v]_2$, as follows:
$$
\stmt_1\OR\stmt[v]_1 \eqgran \stmt_2\OR\stmt[v]_2,\ \stmt_2\OR\stmt[v]_1\eqgran\stmt_3\OR\stmt[v]_2
$$
Without further axioms, it may not be possible to put this together to get transitivity (i.e. to find that $\stmt_1\equ\stmt_3$).

\begin{prop}
The relation $\equ$ is an equivalence relation.
\end{prop}
\begin{proof}
Reflexivity and symmetry are immediate since $\eqgran$ is an equivalence relation. For transitivity, suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$. Let $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$ be infinitesimals with respect to $\stmt[u]$, such that $\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2$ and $\stmt_2\OR\stmt[w]_2\eqgran\stmt_3\OR\stmt[w]_3$. Then we have:
$$
\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
$$
Since $\stmt[v]_1\OR\stmt[w]_2$ and $\stmt[w]_3\OR\stmt[v]_2$ are infinitesimal with respect to $\stmt[u]$ (being finite disjuncts of infinitesimals), we have $\stmt_1\equ\stmt_3$. 
\end{proof}

\begin{defn}\label{smalleru}
Let $\stmt_1, \stmt_2 \in \tdomain$. Then $\stmt_1$ is \textbf{smaller than} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \lequ \stmt_2$, if $\stmt_1 \finer \stmt_2$ or $\stmt_1 \equ \stmt_2$. We define $\stmt_1\leu\stmt_2$ by $\stmt_1\lequ\stmt_2$ and not $\stmt_1\equ\stmt_2$. 
\end{defn}
{\bf Warning:} Transitivity doesn't seem to work as stated (see the proof attempt below). We propose the following modification: say $\stmt_1\lequ\stmt_2$ if there exist statements $\stmt_1'$ and $\stmt_2'$ with $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ such that $\stmt_1'\finer\stmt_2'$. Note that {\bf by totality}, it is sufficient to just say $\stmt_1\finer\stmt_2'$ (no need for $\stmt_1'$).

 \begin{prop}
$\stmt_1 \leu \stmt_2$ if for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$. 
\end{prop}
\begin{proof}
Since $\impossibility$ is infinitesimal, we have $\stmt_1 = \stmt_1\OR\impossibility \sfiner \stmt_2$, and so by Definition \ref{smalleru}, we have that $\stmt_1\lequ\stmt_2$. Now, we must show it is not the case that $\stmt_1\equ \stmt_2$.

Let $\stmt[v]_1,\stmt[v]_2\in\tdomain$ be infinitesimals. Then $\stmt_1\OR\stmt[v]_1 \sfiner \stmt_2 \finer\stmt_2\OR\stmt[v]_2$. We conclude that it cannot be the case that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$ for any infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_1,\stmt[v]_2$. 
\end{proof}

We next show that $\lequ$ is a pre-order, and that it is a partial order with respect to $\equ$-equivalence classes.

\begin{prop}
The relation $\lequ$ is reflexive and transitive.  
\end{prop}
\begin{proof}
Reflexivity is immediate since for any $\stmt\in\tdomain$, we have $\stmt\finer\stmt$, so $\stmt\lequ\stmt$. (This works for both the new and old definition)

Transitivity requires a little more work. Let $\stmt_1,\stmt_2,\stmt_3\in\tdomain$ with $\stmt_1\lequ\stmt_2\lequ\stmt_3$.

{\bf Using the original definition:}
We have several cases to consider.

\emph{Case 1:} $\stmt_1\finer\stmt_2\finer\stmt_3$. Then $\stmt_1\finer\stmt_3$ and so $\stmt_1\lequ\stmt_3$.

\emph{Case 2:} $\stmt_1\equ\stmt_2\equ\stmt_3$. Then $\stmt_1\equ\stmt_3$ (since $\equ$ is an equivalence relation), and so $\stmt_1\lequ\stmt_3$.

\emph{Case 3:} $\stmt_1\equ\stmt_2\finer\stmt_3$. Then for some infinitesimals $\stmt[v]_1$ and $\stmt[v]_2$ (with respect to $\stmt[u]$), we have:
$$
\stmt_1\finer\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2\finer\stmt_3\OR\stmt[v]_2
$$
Unclear what to do here -- need $\stmt_3$ alone on the right-hand side somehow. Case 4 has the same problem. Let's try it with the modified definition (in the ``warning").

{\bf Using the new definition:}
Let $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ with $\stmt_1'\finer\stmt_2'$, and let $\stmt_2''\equ\stmt_2$ and $\stmt_3''\equ\stmt_3$ with $\stmt_2''\finer\stmt_3''$. 
Then for an appropriate choice of infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_i,\stmt[w]_i,\stmt[v]_i',\stmt[w]_i'$, $i=1,2,3$, we have:
$$
\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_1'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \finer \stmt_2'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_2''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \finer
$$
$$
\stmt_3''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \eqgran
\stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1'
$$
Since $\stmt_1 \equ \stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3''$ and $\stmt_3 \equ \stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1' $, we have that $\stmt_1\lequ\stmt_3$.

\end{proof}

\begin{prop}
If $\stmt_1\lequ\stmt_2$ and $\stmt_1\gequ \stmt_2$, then $\stmt_1\equ\stmt_2$.
\end{prop}
\begin{proof}
With the original definition, this is immediate, since we would either have $\stmt_1\eqgran\stmt_2$ (which implies $\stmt_1\equ\stmt_2$), or we would already have $\stmt_1\equ\stmt_2$.

With the modified definition, we need a bit more work. 
\end{proof}

Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotone?

% case: possibilities in u: (1) finite: impose counting measure (2) countable: finitely additive but not countably (3) uncountable: countably additive, legitimate measure

\subsection{Collapsing Orders}

Searching around did not yield the theorems desired, so we will discuss the tools needed here. It turns out that it is not so straightforward to collapse a poset by an equivalence relation. Let's start with a possible definition:

If $(P,\leq)$ is a poset and $\sim$ is a collapsing equivalence relation, then endow the set $P/\sim$ with the relation $\leq/\sim$ defined by $x/\sim \leq/\sim y/\sim$ whenever there exist $x'\in x/\sim$ and $y'\in y/\sim$ with $x'\leq y'$.

It turns out that this is not sufficient. Here is an important example. Let our set be $\{w,x,y,z\}$ and a poset relation defined by $x\leq w$, $y\leq w$, $y\leq z$, with the following Hasse diagram:

\begin{tikzpicture}
    % First, locate each of the nodes and name them
    \node (top) at (0,0) {$w$};
    \node [below left  of=top] (left)  {$x$};
    \node [below right of=top] (bottom right) {$y$};
    \node [above right of=bottom right] (right) {$z$};

    % Now draw the lines:
    \draw [shorten <=-2pt, shorten >=-2pt] (top) -- (left);
    \draw [shorten <=-2pt, shorten >=-2pt] (top) -- (bottom right);
    \draw [shorten <=-2pt, shorten >=-2pt] (bottom right) -- (right);
\end{tikzpicture}

Let's consider two possible cases. Suppose first $w\sim y$, and $x$ and $z$ are only $\sim$ with themselves. Then in order to collapse $\leq$ under $\sim$, we will need $x \leq w\sim y \leq z$ for transitivity.

Consider next the case that $w\sim y$ and $x\sim z$. In order to collapse this, we will need that $w\sim x\sim y \sim z$.

These examples show that not every equivalence relation can collapse a poset into a poset (even if it only collapses a chain).

Note that neither of the above problems are possible with {\bf totality}. It seems that what makes the most sense is collapsing by some equivalence relation, and then extending both the ordering and the equivalence relation to generate a legitimate poset. We can explore this later if needed.

\section{Possible example}

Let $\mathbb{N}$ be the set of natural numbers. We define the sets $D_{i,k} = \{ x \in \mathbb{N} \, | \, x \mod i = k \}$ with $k < i$. Therefore $D_{1,0} = \mathbb{N}$, $D_{2,0} = \{ 2, 4, 6, ... \}$, $D_{2,1} = \{ 1, 3, 5, ... \}$ and so on. Additionally, we say $D_{i,j} \finer D_{k,l}$ for all $k \leq i$ and for all $j$ and $l$. As needed, we may overload the notation by letting $D_{i,j} =  D_{i, j \mod i}$ where $j\mod i$ is the smallest positive residue of $j$ under $i$.

Let $D_{i,j}$ and $D_{k,l}$ be two of these subsets. Consider the intersection $D_{i,j}\cap D_{k,l}$ and suppose it is nonempty; let $n$ be in this intersection. Then $D_{i,j} = D_{i,n}$ and $D_{k,l} = D_{k,n}$, so the intersection is given by $D_{i,n}\cap D_{j,n} = \{x\in \mathbb{N} \, | \, x \mod \gcd(i,j) = n\} = D_{\gcd(i,j),n}$.

Note also that $D_{i,j}^C = \cup_{k=1,k\neq j}^{i-1}D_{i,k}$.

It follows that any finite combination of unions, intersections, and complements of sets of the form $D_{i,j}$ will be empty, or countably infinite.

Next, let $n\in\mathbb{N}$. Consider the union $S_n = \cup_{i=n+1}^{\infty}D_{i,0} = \{n+1,n+2,\ldots\}$, which may be generated by a countable union. Thus $S_n^C = \{1,2,\ldots,n\}$. By combining complements and intersections, we can obtain $\{n\} = S_{n-1}\cap S_n^C$. We can then generate arbitrary subsets of $\mathbb{N}$ using countable operations of union, intersection, and complement on the sets $D_{i,j}$.

Claim: all finite operations of union, intersection, and complement of $D_{i,j}$'s yield sets that may be written as finite unions of $D_{i,j}$'s. Countable operations yield arbitrary subsets of $\mathbb{N}$.

Possible source to read from: http://www.stat.cmu.edu/tr/tr814/tr814.pdf
Note in abstract, gives $1/m$ probability (measure) to $D_{m,j}$ for each $j$. 
\section{Other notes}

Fineness may also capture the concept of physical dimension. In fact, two description of the same units are ``finitely comparable'' in the sense that one gives a finer description of the other of a finite factor. Descriptions of different unites are either ``infinitely comparable'' (i.e. areas are always bigger than lengths) or not comparable (i.e. position and momentum). Take phase space, points should be comparables and in fact should be equigranular $\eqgran$ so that we can compare sets of finite points. Areas are also comparable to each other, and are comparable to points. However, vertical lines (i.e. ranges in momentum alone) are not comparable to horizontal lines (i.e. ranges in position alone). Symplectic geometry, in fact, gives a size to areas and not to lines.

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

### 2021-02-22 Articles/2020-MeasureTheoryFoundation/measureTheoryFoundation.tex
Checkpointing article for measure theory

--- 2021-02-22 Articles/2020-MeasureTheoryFoundation/main.tex
@@
 This working draft summarizes the status of our work on a new approach to the foundations of geometry and measure theory and is intended to gather early feedback.
-The goal is to rederive the usual concepts in geometry, measure theory and probability theory from a simpler more general structure, namely a preorder.
-This construction will force us to make explicit the physical requirements that are baked into the use of the common concepts and provide us with a framework that is, hopefully, still valid when those assumptions fail.
-
-The general idea is similar to what is done in comparative probability, yet we want the techniques extend from a single measure to a family of measures that allow us to define and compare sizes of sets at different scales (e.g.
+The goal is to rederive the usual concepts in geometry, measure theory and probability theory from a simpler more general structure.
+This construction will force us to make explicit the physical requirement that are baked into the use of the common concepts and provide us with a framework that is, hopefully, still valid when those assumptions fail.
+
+The general idea is similar to what is done in comparative probability, yet the techniques extends from a single measure to a family of measure that allow us to define and compare sizes of sets at different scales (e.g.
 points, lines, surfaces, ...).
@@
     \item \statement{The position of the object is between 0 and 1 meters} $\narrower$ \statement{The position of the object is between 0 and 1 kilometers}
-    \item \statement{The fair die landed on 1} $\narrower$ \statement{The fair die landed on 1 or 2}
-    \item \statement{The first bit is 0 and the second bit is 1} $\narrower$ \statement{The first bit is 0}
+    \item \statement{The fair dice landed on 1} $\narrower$ \statement{The fair dice landed on 1 or 2}
+    \item \statement{The first bit is 0 and the second bit is 2} $\narrower$ \statement{The first bit is 0}
 \end{itemize}
-In these cases, the first statement has to be ``contained'' in the second, which is more general.
+In these cases, the first statements has to be ``contained'' in the second, which is more general.
 
@@
     \item \statement{The position of the object is between 0 and 1 meters} $\finer$ \statement{The position of the object is between 2 and 3 kilometers}
-    \item \statement{The fair die landed on 1} $\finer$ \statement{The fair die landed on 3 or 4}
-    \item \statement{The first bit is 0 and the second bit is 1} $\finer$ \statement{The third bit is 0}
+    \item \statement{The fair dice landed on 1} $\finer$ \statement{The fair dice landed on 3 or 4}
+    \item \statement{The first bit is 0 and the second bit is 2} $\finer$ \statement{The third bit is 0}
 \end{itemize}
-In these cases, the first statement may not be contained or overlap with the second.
+In these case, the first statement may not be contained or overlap with the second.
 Even though the statements come from different disciplines (i.e.
@@
 Therefore the notion of fineness is linked to the notion of ``counting'' of the possible outcomes.
-It should be no surprise, then, that the same concept is compatible with the different disciplines above: geometrical size quantifies the ways we can arrange objects; probability theory quantifies the ratio of expected cases; information theory needs to quantify the different possible messages stored, computed or transmitted through an information system.
+It is should be no surprise, then, that the same concept is compatible with the different disciplines above: geometrical size quantifies the ways we can arrange objects; probability theory quantifies the ratio of expected case; information theory needs to quantify the different possible messages stored, computed or transmitted through an information system.
 Therefore the common structure we propose is not simply a mathematical pattern among disconnected structures: it is a single conceptual framework.
@@
 \begin{itemize}
-    \item $\stmt_1$ = \statement{The horizontal position of the object is exactly 0 meters}
+    \item $\stmt_1$ = \statement{The horizontal position of the object is exactly 0 meter}
     \item $\stmt_2$ = \statement{The horizontal position of the object is exactly 1 or 2 meters}
@@
 \end{itemize}
-Therefore we need to construct a family of measures, each identified by picking a particular statement that acts as a reference, as a unit.
-
-On the other hand, geometric structures do provide measure for different dimensionality by starting from a lowest definition (i.e.
-line distance for Riemannian and surface are for symplectic) and then constructing higher dimensional measures through products.
-While want to recover those cases as well, we want to be able to capture the physical idea of a unit system, which is not typically done.
-In that case, there will be dimensions that are not directly comparable (e.g.
-pressure, volume, number of particles, chemical potential) but are still comparable under appropriate composition (i.e.
-pressure times volume with number of particles times chemical potential).
+Therefore we need to construct a family of measures, each identified by picking a particular statements that acts as a reference, as a unit.
+This also allows to define more precisely how the physical idea of a unit system can be captured mathematically, something that has never been addressed consistently.
 
@@
 
-In this overview we give a more detailed account of our goal and the status of our search.
-This will give a general idea of the mathematical issues, but it will also focus on the construction at a physical/conceptual level.
+We start by giving an overview of the approach.
+We will give a summary of the main steps and the results, but we will also discuss and justify the starting points and the construction at a physical/conceptual level.
 It is important for our larger project that the mathematical detail proceeds in lock step with the conceptual and physical understanding.
-Later sections, instead, will concentrate on the bare formal aspects.
-
-\subsection{Fineness and granularity}
+This will allow us in the later section to concentrate on the full  mathematical detail.
 
@@
 Conceptually, this represents all the statements that can be associated with an experimental verification procedure.
-The full detail of why this is the correct abstraction is discussed in detail in our previous work.\footnote{See \url{https://assumptionsofphysics.org/book/}} It is a property of this type of spaces that each statement $\stmt \in \tdomain$ can be expressed as a disjunction of possible cases, of the  \textbf{possibilities} of the domain.
-Formally, each possibility is a minterm\footnote{A minterm is a disjunction in which each term appears once, either negated or not.} of the countable basis.
-The theoretical domain can then equivalently be represented by a $\sigma$-algebra.\footnote{This conceptually maps perfectly to what one has in probability: a sample space (i.e.
+The full detail of why this is the correct abstraction is discussed in detail in our previous work.[TODO: cite book] It is a property of this type of spaces that each statement $\stmt \in \tdomain$ can be expressed as a disjunction of possible cases, of the  \textbf{possibilities} of the domain.
+Formally, each possibility is minterm of the countable basis.\footnote{The theoretical domain can then be re-expressed as a $\sigma$-algebra one would use in probability theory: a sample space (i.e.
 the set of possibilities) and a set of events (i.e.
@@
 We call \textbf{narrowness} this partial order as it describes whether a statement gives a narrower description of another as we saw in the introduction.
-The narrowest statement in the domain is the impossibility $\impossibility$ and the broadest one is the certainty $\certainty$.\footnote{In terms of the $\sigma$-algebra, narrowness corresponds to set inclusion, the certainty is the full set and the impossibility is the empty set.}
-
-Conceptually, narrowness allows us to say that a description is more refined than another only if the first is fully contained in the second.
+The narrowest statement in the domain is the impossibility $\impossibility$ and the broadest one is the certainty $\certainty$.
+
+Conceptually, narrowness allows us to say that a description is more refined than another only if it is fully contained by another.
 It cannot compare descriptions that do not overlap, that are incompatible.
-That is why it is not sufficient to quantify the granularity, the level of detail, of the description.
-The ability to compare the granularity of the description of two statements, therefore, is captured by an additional binary relationship $\finer$ we call \textbf{fineness}.\footnote{In terms of the $\sigma$-algebra, fineness allows us to say that one set is ``bigger'' than another.}
-
-The whole program can be summarized in the following question: what intuitive conceptual properties does fineness have physically that, once properly formalized mathematically, allow us to recover measures, probability and geometrical structures?\footnote{In terms of the $\sigma$-algebra, how do we go from being able to compare the size of sets to being able to quantify it?}
-
-\subsubsection{Necessary properties for fineness}
-
-The first set of properties are the ones we must expect fineness to have in all cases.
-At this point, there are three, identified by DeFinetti for comparative probability, that we believe are necessary and that work as simple and intuitive starting points.
-
-\textbf{Boundness}.
-As we mentioned in the introduction, a description is finer than another if it can be true in fewer cases.
-This intuitive notion is in general ill-defined except for two cases.
-A certainty $\certainty$ (e.g.
+That is why it is not sufficient to quantify the level of description.
+The ability to compare the granularity of the description of two statements, therefore, is captured by an additional binary relationship $\finer$ we call \textbf{fineness}.
+To properly capture this idea, finesses must have at least three property.
+
+As we saw, a description is finer than another if it can be true in fewer cases.
+This intuitive notion is in general ill-defined except than in two cases.
+A certainties $\certainty$ (e.g.
 \statement{the position is between minus infinity and plus infinity}) is true in all cases, therefore it must be the coarsest statement.
 An impossibility (e.g.
-\statement{the position is between 0 and 1 and also between 2 and 3}) is false in all cases, therefore it must be the finest statement.
+\statement{the position is between 0 and 1 and also between 2 and 3}) is false in all all cases, therefore it must be the finest statement.
 This means that fineness must be uniquely bound by the impossibility and the certainty.
 Any contingent statement must be strictly between the two since there are some cases in which it will be true and others in which it will be false.
-This is the first required property.\footnote{In terms of the $\sigma$-algebra, the full set is strictly the biggest set and the empty set is strictly the smallest set.}
-
-\textbf{Transitivity}.
+This is the first required property.
+
 Additionally, if $\stmt_1$ gives finer description than $\stmt_2$ and $\stmt_2$ gives a finer description than $\stmt_3$ then it must be that $\stmt_1$ gives a finer description than $\stmt_3$.
@@
 
-\textbf{Offset monotonicity}.
-Lastly, if we take two descriptions and we make them less precise ``in the same way'', we must have the same relationship in fineness.
-More precisely, suppose we have two descriptions $\stmt_1$ and $\stmt_2$ and we pick a third description $\stmt$ that is incompatible\footnote{Two statements are incompatible if they can't be both true at the same time: their conjunction $\stmt_1 \AND \stmt \equiv \impossibility$ is impossible.
-This is the analogue of disjoint sets in a $\sigma$-algebra.} with both.
+Lastly, fineness must be related to narrowness: if we have a description that is strictly narrower than another, we expect it to be strictly finer as well.
+It turns out\cite{definetti} that a single property is enough to specify the relationship.
+Suppose we have two descriptions $\stmt_1$ and $\stmt_2$ and we pick a third description $\stmt$ that is incompatible with both.
+By incompatible we mean that two statements can't be both true at the same time (i.e.
+their conjunction $\stmt_1 \AND \stmt \equiv \impossibility$ is impossible).
 The respective disjunctions $\stmt_1 \OR \stmt$ and $\stmt_2 \OR \stmt$ will be respectively ``equally'' broader than the respective original statement.
 Therefore we will want them to be ``equally'' coarser as well.
-In other words, it must be that $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \OR \stmt \finer \stmt_2 \OR \stmt$: fineness must be offset monotonic\footnote{Some authors call this property ``monotony'', which we don't find descriptive.
-We prefer ``offset monotonicity'' as it makes it clear that adding or removing a ``constant'' on both sides is a monotonic operation.} in the sense that it is invariant if we add or remove a statement from both sides, and this is the third and last required property.\footnote{In terms of the $\sigma$-algebra, a set is smaller than another if and only if it is also smaller when we increase both sets with another disjoint from the two.}
-
-From these three properties, we can show the following:
-\begin{description}
-    \item[Preorder.] Fineness is a preorder over the theoretical domain $\tdomain$.
-    \item[Monotonicity.] Fineness respects narrowness: $\stmt_1 \narrower \stmt_2$ implies $\stmt_1 \finer \stmt_2$.
-    \item[Equigranularity is an equivalence.] Equigranularity, the symmetrization of fineness, is an equivalence class, and fineness is an order of those equivalence classes.
-That is, fineness orders the levels of description.
-\end{description}
-
-While we are not convinced these are sufficient properties to define fineness, we are not aware of other properties that would make the cut.
-
-\subsubsection{Optional properties for fineness}
-
-Optional properties for fineness include properties that are desirable, but are known to fail in some cases. These will be properties of a particular domain, rather than fineness itself.
-
-\textbf{Equigranularity of possibilities}.
-Possibilities provide a complete characterization of a domain: if we know one possibility to be true (e.g.
-\statement{the mass of the photon is exactly 0 eV}), we know the truth value of all other statements (e.g.
-\statement{the mass of the photon is between x and y eV}).
-Therefore it is natural to assume that all possibilities are equigranular.
-In fact, this is a necessary property if we want to have a unique measure when we take a possibility as a unit.\footnote{The measure, in this case, is simply the counting measure and equigranularity of the possibilities plays a role similar to the principle of indifference in classical probability} However, this property can be easily violated by creating a subdomain.
-
+Therefore it must be that $\stmt_1 \finer \stmt$ if and only if $\stmt_2 \finer \stmt$.
+In other words, fineness must be invariant under addition\footnote{Some authors call this property ``monotony'', which we don't find descriptive.
+In measure theory, union of disjoint sets plays the role of addition and, in our framework, disjunction of incompatible statements plays the same role.} and this is the third and last required property.
+
+From these starting points one can define and prove a series of useful properties.
+Most importantly, that fineness is a preorder and that it is monotonic with respect to narrowness.
+However, these properties will not be enough to recover a unique quantifiable concept of size.
+To do so, we will require two additional properties which are not strictly a property of fineness itself but a property of the domain.
+That is, fineness happen to have that property for that specific set of statements but it is not in general true.
+
+We say a theoretical domain $\tdomain$ is uniform if all possibilities are equigranular: they provide the same level of description.\footnote{This plays a role similar to the principle of indifference in classical probability, but does not suffer the same problems.} This is cannot be true for all domains since we can create domains whose possibilities are less refined.
 Suppose we have a uniform domain with three possibilities: red, green, blue.
-We can take a ``color blind'' subdomain that cannot distinguish between red and green.
+We can take a ``color bline'' subdomain that cannot distinguish between red an green.
 This would have two possibilities: red/green and blue.
 This will not be uniform.
-The idea is that if the domain is not uniform, the cases are not really describing ``the same thing'': some information is missing.
-Equigranularity of possibilities, then, is not a property of fineness itself, but a property of a domain that is ``well-posed''.
-
-\textbf{Totality}.
-If our objective is to construct a measure, totality is needed in the sense that every statement (i.e.
-every Borel set of possibilities) needs to be comparable with every other if we are to assign a unique number.
-However, statements whose size is defined in units of different physical dimensions (i.e.
-length, pressure, mass) should not be comparable.
-Lack of totality, in fact, should be a feature that allows us to formally construct a unit system.
-
-Another counter example come from infinite ranges.
-Take the real line and consider the set of all positive numbers and the set of all negative numbers: should they be comparable?
-If we say they are equal, then that would tell us that there are more cases before 1 than after.
-This means that the space is not symmetrical under translation.
-If the real values represent distance between points, this is bad as it violates the idea that space is uniform.
-If the real values represents charge, however, this would actually make sense: the idea that there are exactly the same values for positive and negative charges does make sense, and the symmetry we are imposing is natural since zero is really a special number.
-Therefore different choices of comparability for the same Borel algebra is actually a feature we do not want to lose in general.
-
-However, once we pick the unit, we should be able to find the ``right'' set of statements on which to create the measure.
-One approach would be to ask the ``largest set of statements mutually comparable in fineness that contain the unit''.
-In general, we have no expectation that this would be unique.
-So one question would be: when is this true?
-
-\subsubsection{Unsettled properties for fineness}
-
-\textbf{Boolean lattice}.
-Since equigranularity classes form a bounded partially ordered set, one can ask whether they should form a lattice.
-Physically this would make sense: we would be saying that given a set of statements $S \subseteq \tdomain$, we can always find a level of description that is the finest among all coarser descriptions than all the statements (i.e.
-$\bigveedot S$) and a level of description that is the coarsest among all finest descriptions than all the statements (i.e.
-$\bigwedgedot S$).
-However, it is not clear whether this is required, something only to be imposed in some cases or something that can be already proven.
-
-Regardless, the map from a statement to its level of description (i.e.
-equivalence class in equigranularity), while it is an order isomorphism and preserves the top and bottom, is not a lattice isomorphism.
-That is, it does not preserve the meet and join.
-Let us note $\dot{\stmt} = {\stmt}_{/\eqgran}$ the level of granularity of $\stmt$.
-Now suppose that $\stmt_1 \eqgran \stmt_2 \nequiv \impossibility$ and $\stmt_1 \ncomp \stmt_2$.
-Then $\dot{\stmt}_1 = \dot{\stmt}_2 \neq \dot{\impossibility}$.
-Now suppose that $\stmt \to \dot{\stmt}$ is a lattice isomorphism.
-Then $\dot{\impossibility} = (\stmt_1 \AND \stmt_2)_{/\eqgran} = \dot{\stmt}_1 \wedgedot \dot{\stmt}_2 = \dot{\stmt}_1 \wedgedot \dot{\stmt}_1 = \dot{\stmt}_1$, which contradicts the premise.
-
-\textbf{Addition}.
-We also need to understand how to define the addition and subtraction.
-Conceptually, the addition of two level of precision $\dot{\stmt}_1 + \dot{\stmt}_2$ means returning the level of precision that can be achieved by taking the disjunction $\stmt_1 \OR \stmt_2$ of two incompatible statements $\stmt_1 \in \dot{\stmt}_1$ and $\stmt_2 \in \dot{\stmt}_2$.
-This, of course, is not possible if the statements are ``bigger'' than ``half'' the space.
-A more order theoretical definition would be to take all possible disjuctions of $\stmt_1 \OR \stmt_2$ such that $\stmt_1 \in \dot{\stmt}_1$ and $\stmt_2 \in \dot{\stmt}_2$ and return the coarsest level of description.
-That is, the sum is the least upper bound in fineness of all possible disjunctions.
-One should be able to prove that the incompatible disjunction corresponds to that level of description.
-
-It is not clear whether the axioms above are enough to show that the addition always exists.
-It may be the existence of the addition, or a premise that leads to it, needs to be added.
-It is not clear whether, physically, such addition always makes sense, particularly in the context of incomparable descriptions (e.g.
-descriptions quantified in different units).
-
-\textbf{Product}.
-Another thing that needs to be defined/recovered is that independent domains lead to the product of the measure.
-This means that there must be some compatibility condition between statement independence and fineness.
-It is not clear whether this compatibility is already embedded in the given axioms or not.
-The issue is that, in comparative probability, independence is only defined later in terms of the probability measure.
-In our framework, we have already a notion of independence that is defined on statements and logic operation, which is based on the idea of whether fixing the truth value of some statements will also fix the truth value of other statements.
-
-\subsection{A family of measures}
-
-The main goal is to understand under what conditions we can pick a statement $\stmt[u]$ and find a unique measure $\mu_{\stmt[u]} : \tdomain \to \mathbb{R}$ with the following properties:
-\begin{description}
-    \item[Normalization.] The value for the unit is one,  $\mu_{\stmt[u]}(\stmt[u]) = 1$.
-    \item[Monotonicity.] It is a monotonic function with respect to fineness; if $\stmt_1 \finer \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
-    \item[Finite additivity.] If $\stmt_1 \ncomp \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$.
-\end{description}
-
-\textbf{Countable additivity}.
-Additionally, we want to understand when finite additivity can be extended to countable additivity.
-In general, countable additivity will fail.
-Consider a space with countably many possibilities, all of them equigranular.
-If we take the certainty as a unit, each possibility is infinitely smaller, and therefore it must be assigned a measure zero.
-Therefore, the sum (i.e.
-the disjunction) of infinitely many measure zero statements is one.
-
-\textbf{Equigranular possibilities}.
-If a possibility is taken as a unit, then the equigranularity of possibilities is a sufficient condition and it leads to the counting measure on the possibilities.
-This is strictly not necessary: we could have a domain with 5 possibilities $\{x_1, x_2, x_3, y_1, y_2\}$ such that $x_1 \eqgran x_2 \eqgran x_3$, $y_1 \eqgran y_2$ and $x_1 \OR x_2 \OR x_3 \eqgran y_1 \OR y_2$.
-If we took $x_1$ as a unit, this would lead to a unique measure such that $\mu(x_1)=\mu(x_2)=\mu(x_3)=1$ and $\mu(y_1)=\mu(y_2)=3/2$.
-This also tells us that we may not be really interested in necessary conditions, since these cases are contrived.
-Rather, we are interested in sufficient conditions that can be well justified from the physics.
-
-\textbf{Lack of reverse monotonicity}.
-Monotonicity only works in one direction.
-If a statement describing an area is taken as unit, all statements about lines, points and countable sets of both are going to be measure zero.
-Therefore equal measure does not mean equal fineness.
-This is why fineness can generate a family of measures: it has more information than a single one.
-When we pick a unit, then, the measure has to ``collapse`` the equivalence classes of equigranularity appropriately.
-This is tightly linked with a notion of infinitesimal (i.e.
-the ordering over the level of granularity is non-archemedian, while the real numbers are).
-The idea is to proceed in the following way:
+The idea is that a uniform domain is a what we want since all cases give us the same level of description.
+If that's not the case, it's because we are ``missing'' something and we would need to find out what.
+Uniformity is the first additional property.
+
+We say a theoretical domain $\tdomain$ is totally comparable, or total, if fineness is a total preorder: given any two statements, one is always finer than the other.
+While this may seem like a reasonable and desirable property, it is actually not always the case.
+There is no sense, for example, that a range of positions is bigger or smaller than a range of mass or energy.
+On a symplectic manifold, for example, we can compare areas to each other, yet we cannot compare lengths.
+In general, the fact that fineness is not total may actually allow us to properly define mathematically the idea of physical dimension.
+However, for symplicity, in this paper we will concentrate on domains that are totally comparable, leaving the other case for future work.
+
+At this point, we pick a unit $\stmt[u]$ and attempt to construct a unique measure on the sub-algebra of all statements narrower than $\stmt[u]$.
+One problem is that $\tdomain$ is not in general Archemedian: we are going to have statements that are ``infinitely smaller'' than $\stmt[u]$.
+In [XXX] we define what it means for a statement to be infinitesimal with respect to another.
+We then define an equivalence relationship $\equ$ such that two statements are equal in $\stmt[u]$ if they are equigranular up to an infinitesimal of $\stmt[u]$.
+We then collapse the fineness preorder with that equivalence relationship, and obtain a new preorder $\lequ$.
+At this point we can show that there exists a unique measure $\mu_{\stmt[u]}$ such that $\mu_{\stmt[u]}(\stmt[u])=1$ and $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$ if and only if  $\stmt_1 \lequ \stmt_2$.
+Since we are restricting ourselves to all statements narrower than $\stmt[u]$, the measure we find is a measure bounded at one, which means we can essentially reuse the techniques from comparative probability even though the context is different.
+
 \begin{enumerate}
-    \item pick a unit $\stmt[u]$ and define the notion of infinitesimal
-    \item define the preorder $\lequ$ and the equivalence $\equ$ that compares levels of description up to infinitesimals
-    \item show that $\lequ$ is order isomorphic to the reals
-    \item show that we can pick a particular isomorphism that respects linear addittivity.
-\end{enumerate}
-
-\subsubsection{Infinitesimals and finite comparison}
-
-The Archemedian property is typically discussed in terms of linearly ordered groups.
-\footnote{See \url{https://en.wikipedia.org/wiki/Archimedean_property} for an introduction.} Roughly speaking, if we pick an element $y$ an element $x$ is infinitesimal with respect to $y$ if $\sum\limits_{i=1}^n x < y$ for any $n$.
-The group is Archemedian if no element is infinitesimal with respect to the other.
-We need to extend this notion to our ordering in fineness.
-
-The first issue, then, is to find the correct definition to extend this concept on ordering.
-The addition discussed about (i.e.
-roughly the disjoint union of statements) should play the role of the group operation.
-Conceptually, a statement $\stmt$ is infinitesimal with respect to another $\stmt[u]$ if for any $n$ we can find a finite set of equigranular incompatible statements whose disjunction is finer than $\stmt[u]$.
-That is, we need infinitely many statements equigranular to $\stmt$ to cover a statement of the same granularity than $\stmt[u]$.
-
-The main issue here is that, since the ordering is potentially not total, it is not clear how one can prove the existence of said sets of statements even in the simplest cases.
-
-Assuming those issues are solved, infinitesimals come in two types: countable and uncountable.
-A countable infinitesimal will be such that the disjoint conjunction of countably many equigranular statements will be bigger than the unit.
-This is the case where, in the end, the measure will only be finitely additive: infinitely many zero measure elements will some to a finite element.
-If all infinitesimal with respect to the unit are uncountable, instead, countably many zero measure elements will still form a zero measure element.
-
-This insight may give us a way to define ``finite'' and ``infinite'' elements in an absolute way.
-A finite element has no countable infinitesimal.
-For example, take the unit interval over the reals and consider the standard Lebesque measure over the Borel sets (which would correspond to our theoretical domain).
-All sets of measure zero are uncountable infinitesimals with respect to the unit: since the measure is countably additive, any disjoint union of countably many measure zero sets is going to be measure zero, therefore it will not be as big as the unit.
-Natuarally, if we take an uncountably many sets (i.e.
-all the singletons) we would cover the whole unit.
-Conversely, the full real line can be covered by countably many sets of unit size.
-While the unit interval is infinitesimally small with the respect with the whole line, it is a countable infinitesimal.
-This definition of ``finiteness'' would map to physical requirements much better than, for example, the topological notion of compactness.
-
-Once a notion of infinitesimals is well defined, we would define a preorder $\lequ$ and an equivalence relationship $\equ$ that compares the level of description of the statements up to infinitesimals.
-
-\subsubsection{Recovering bounded measures}
-
-The next step is to understand how the properties of $\lequ$ would compare to the requirements set by, for example Villegas \cite{villegas}.
-In the context of comparative probability, there are theorems that allow to go from a preorder to a probability measure.
-The idea would be to reuse, as much as possible, the ideas from that literature.
-Once we pick a unit $\stmt[u]$, we can restrict ourselves to all statements narrower than $\stmt[u]$.
-At that point, we are looking for has all the mathematical properties of a probability measure as the biggest element, $\stmt[u]$, has to have measure one.
-
-One issue will be the problem of countable infinitesimals, which are incompatiable with a countable measure.
-Therefore one of the assumptions in comparative probability will effectively exclude the case.
-We need to understand how that works, so that we can adapt the proofs and allow for only finitely additive measures.
-
-The other issue is that the measure will not be, in general, on the whole domain $\tdomain$, but on a subset $\tdomain_{\stmt[u]}$ that must be totally comparable with $\lequ$.
-It is not clear how to recover this set.
-It may be that the equivalence classes of $\equ$ will make our job easier: if two infinitesimal are not comparable in fineness, they will still be merged together.
-Therefore  incomparable statements may be comparable in $\lequ$ if they can be compared up to an infinitesimal.
-
-The goal here is to solve the above problems to be find the prerequisites for the following proposition: given a statement $\stmt[u] \in \tdomain$ we can define a maximal set of mutually comparable statements in $\lequ$ that contains $\stmt[u]$, upon which we can define a unique $\mu_{\stmt[u]} : \tdomain_{\stmt[u]} \to \mathbb{R}$ such that:
-\begin{description}
-    \item[Normalization.] The value for the unit is one,  $\mu_{\stmt[u]}(\stmt[u]) = 1$.
-    \item[Order isomorphism.] $\stmt_1 \lequ \stmt_2$ if and only if $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
-    \item[Finite additivity.] If $\stmt_1 \ncomp \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$.
-\end{description}
-Moreover, if $\stmt[u]$ has no countable infinitesimals, $\mu_{\stmt[u]}$ is countably additive.
-
-\subsubsection{Extending the measures}
-
-Once we know how to construct measures on the sub-algebra of statements narrower than the unit, we should be able to ``stitch them together'' in the following way.
-
-\begin{enumerate}
+    \item Show that if we pick a unit $\stmt[u]$, we can construct a unique measure on the sub-algebra of all statements narrower than $\stmt[u]$.
+The idea is to define an ordering $\leq_{\stmt[u]}$ that ``squashes" all ``small'' statements to the same equivalence class in $=_{\stmt[u]}$ of the impossibility.
+An open problem is to understand how to identify the measure zero sets.
     \item Prove that if $\stmt[v] \narrower \stmt[u]$ then $\mu_{\stmt[u]}(\stmt[w]) = \mu_{\stmt[u]}(\stmt[v]) \mu_{\stmt[v]}(\stmt[w]) $ for all $\stmt[w] \narrower \stmt[v]$.
-To achive this, we will need to prove that the infinitesimals with respect to $\stmt[u]$ are exactly the infinitesimals with resepect to $\stmt[v]$.
+Note that, once this is proven, it tells us that measure zero sets in $\stmt[u]$ are exactly the measure zero sets in $\stmt[v]$.
     \item Given two statements $\stmt[u] \narrower \stmt[v]$, we extend the measure on $\stmt[u]$ by defining $\mu_{\stmt[u]}(\stmt[v]) = 1/ \mu_{\stmt[v]}(\stmt[u])$.
@@
 
-One issue will be, again, proving the consistent of comparability with different units since all the domains $\tdomain_{\stmt[u]}$ do not contain all the statements.
+To recover a measure, the idea is to pick a possible statement $\stmt[u] \in \tdomain$ which serves as a unit and construct a total preorder $\leq_{\stmt[u]}$ over $\tdomain_{\stmt[u]} \subseteq \tdomain$.
+This preorder will induce to a measure.
+That is, we will be able to find a unique $\mu_{\stmt[u]} : \tdomain_{\stmt[u]} \to \mathbb{R}$ is such that:
+\begin{itemize}
+    \item $\mu_{\stmt[u]}(\stmt[u])=1$ 
+    \item $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$ if and only if  $\stmt_1 \leq_{\stmt[u]} \stmt_2$
+    \item given $\stmt_1 \ncomp \stmt_2$ we have $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$
+\end{itemize}
+and .
+
+The total preorder $\leq_{\stmt[u]}$ will need to be compatible with fineness.
+That is, if $\stmt_1 \finer \stmt_2$ then $\stmt_1 \leq_{\stmt[u]} \stmt_2$.
+It will not be true that $\stmt_1 \sfiner \stmt_2$ then $\stmt_1 <_{\stmt[u]} \stmt_2$.
+The converse is not true: there will be statements such that $\stmt_1 \leq_{\stmt[u]} \stmt_2$.
+The totality also needs to be constructed.
+
+Given a total preorder $\leq_{\stmt[u]}$, we say a statement $\stmt$ is zero-comparable to $\stmt[u]$ if and only if $\impossibility =_{\stmt[u]} \stmt$.
+If $\certainty =_{\stmt[u]} \certainty \AND \NOT \stmt[u]$ then a $\stmt$ is infinitely-comparable to $\stmt[u]$ if and only if $\certainty =_{\stmt[u]} \stmt$.
+If $\certainty >_{\stmt[u]} \certainty \AND \NOT \stmt[u]$ then there are no statements infinitely-comparable to $\stmt[u]$.
+All other statements that are comparable are said finitely comparable.
+If two statements $\stmt_1$ and $\stmt_2$ have the property that $\stmt_1\AND\stmt_2\equ\impossibility$, we say that the pair of statements is null-compatible.
+This will be denoted by $\stmt_1\ncomp_{\stmt[u]}\stmt_2$.
+NOTE: I'm not attached to this terminology; it is just a common concept that deserves a name.
 
@@
 
-The fact that the $\sigma$-algebra is generated by a countable set means it is separable. It also means that each element can be constructed as a disjunction of a set of possibilities.
+The fact that the $\sigma$-algebra is generated by a countable set means it is separable.
+
+As a simplifying hypothesis, we assume fineness is a total preorder.
 
@@
 	\begin{itemize}
-		\item boundedness: for each contingent statement $\stmt \in \tdomain$ we have $\impossibility \sfiner \stmt \sfiner \certainty$
+		\item unique boundedness: for each contingent statement $\stmt \in \tdomain$ we have $\impossibility \sfiner \stmt \sfiner \certainty$
 		\item transitivity: if $\stmt_1 \finer \stmt_2$ and $\stmt_2 \finer \stmt_3$ then $\stmt_1 \finer \stmt_3$
-		\item offset monotonicity: if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \finer \stmt_2$ if and only if $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$.
+		\item addition invariance?: if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \finer \stmt_2$ if and only if $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$.
 	\end{itemize}
+\end{defn}
+
+\begin{defn}
+     Let $\stmt_1, \stmt_2 \in \tdomain$. We say $\stmt_1$ is \textbf{equigranular} to $\stmt_2$ (noted $\stmt_1 \eqgran \stmt_2$) if $\stmt_1 \finer \stmt_2$ and $\stmt_2 \finer \stmt_1$.
 \end{defn}
@@
 
-\begin{prop}
-    Fineness is a preorder.
-\end{prop}
-
-\begin{proof}
-    Let $\stmt \in \tdomain$.
-Since $\stmt \narrower \stmt$, by \ref{fineness_properties}.3 $\stmt \finer \stmt$ and therefore fineness is reflexive.
-By \ref{fineness} fineness is also transitive and therefore is a preorder.
-\end{proof}
-
-\begin{defn}
-    We say $\stmt_1$ is \textbf{equigranular} to $\stmt_2$ (noted $\stmt_1 \eqgran \stmt_2$) if $\stmt_1 \finer \stmt_2$ and $\stmt_2 \finer \stmt_1$.
-\end{defn}
-
-\begin{prop}
-Other properties useful properties:
-\begin{enumerate}
-    \item (monotony as in Villegas) if $\stmt[v]_1 \ncomp \stmt[v]_2$, then $\stmt_1 \finer \stmt[v]_1$ and $\stmt_2 \finer \stmt[v]_2$ imply  $\stmt_1 \OR \stmt_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
-    \item if $\stmt_1 \ncomp \stmt[v]_1$, $\stmt_2 \ncomp \stmt[v]_2$ and $\stmt[v]_1 \eqgran \stmt[v]_2$, then $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \OR \stmt[v]_1 \finer \stmt_2 \OR \stmt[v]_2$.
-\end{enumerate}
-\end{prop}
-
-\begin{proof}
-
-\end{proof}
-
-\begin{prop}
-    Equigranularity is an equivalence relationship.
-\end{prop}
-
-\begin{proof}
-    It is an equivalence relationship because it is the symmetrization of a preorder.
-Since fineness is reflexive then equigranularity is also reflexive: for all $\stmt \in \tdomain$ $\stmt \finer \stmt$ and therefore $\stmt \eqgran \stmt$.
-Since fineness is transitive then equigranularity is also transitive: $\stmt_1 \eqgran \stmt_2 \eqgran \stmt_3$ means $\stmt_1 \finer \stmt_2 \finer \stmt_3$ therefore $\stmt_1 \finer \stmt_3$ but also $\stmt_3 \finer \stmt_2 \finer \stmt_1$ therefore $\stmt_3 \finer \stmt_1$ and finally $\stmt_1 \eqgran \stmt_3$.
-Additionally, equigranularity is reflexive by definition and is therefore an equivalence relationship.
-\end{proof}
-
-\begin{defn}
-    Let $\tdomain_{/\eqgran}$ be the granularity equivalence classes. Let $(\cdot)_{/\eqgran} : \tdomain \to \tdomain_{/\eqgran}$.
-\end{defn}
-
-\begin{prop}
-    The function $(\cdot)_{/\eqgran}$ is an order homomorphism.
-\end{prop}
-
-\begin{proof}
-    Let $\stmt_1, \stmt_2 \in \tdomain$.
-Suppose $\stmt_1 \narrower \stmt_2$ then by \ref{fineness_properties}.3 $\stmt_1 \finer \stmt_2$ and therefore ${\stmt_1}_{/\eqgran} \finer {\stmt_2}_{/\eqgran}$.
-Then $(\cdot)_{/\eqgran}$ is an order isomorphism.
-\end{proof}
-
-\begin{remark}
-    Regardless of whether $\tdomain_{/\eqgran}$ is a lattice, $(\cdot)_{/\eqgran}$ can never be a lattice homomorphism.
-It will be crucial that two equigranular statements can be incompatible, and this in contradiction with $(\cdot)_{/\eqgran}$ being a lattice homomorphism.
-In fact, suppose $\tdomain_{/\eqgran}$ is a lattice and let $\wedgedot$ be its meet.
-Suppose that $\stmt_1 \eqgran \stmt_2 \nequiv \impossibility$ and $\stmt_1 \ncomp \stmt_2$.
-Then ${\stmt_1}_{/\eqgran} = {\stmt_2}_{/\eqgran} \neq \impossibility_{/\eqgran}$.
-Now suppose that $(\cdot)_{/\eqgran}$ is a homomorphism.
-Then $\impossibility_{/\eqgran} = {(\stmt_1 \AND \stmt_2)}_{/\eqgran} = {\stmt_1}_{/\eqgran} \wedgedot {\stmt_2}_{/\eqgran} = {\stmt_1}_{/\eqgran} \wedgedot {\stmt_1}_{/\eqgran} = {\stmt_1}_{/\eqgran}$, which contradicts the premise.
-\end{remark}
-
-\begin{remark}
-We claim there exist domains which have all possibilities equigranular but fineness is not a total order.
-Suppose we have a theoretical domain in bijection with the power set of the natural numbers, $\mathcal{P}(\mathbb{N})$.
-The partial order of narrowness may be given by the ordering of the power set by set containment, and the possibilities (which also serve as a countable generating set) are given by $\mathbb{N}$.
-The hypotheses of the fineness relation (Definition \ref{fineness}) are satisfied if we order contingent statements by cardinality, with the additional requirement that infinite strict subsets of $\mathbb{N}$ are considered strictly finer than the certainty $\mathbb{N}\in\mathcal{P}(\mathbb{N})$.
-All finite sets will be totally (pre)ordered, but infinite sets will not be ordered.
-We note that one could just say that $\stmt_1\finer\stmt_2$ for every pair of infinite contingent statements $\stmt_1,\stmt_2$, but even if we remove those from the relation $\finer$, the axioms are still satisfied, and the (pre)order is not total.
-\end{remark}
-
 \begin{defn}
@@
 
-\begin{defn}
-Let $(S,\leq)$ be a poset.
-A \textbf{chain} is a subset $T\subseteq S$ which is totally ordered (i.e.
-every pair of elements in $T$ is comparable in $\leq$).
-A chain is said to be \textbf{maximal} if it is not a proper subset of any other chain, i.e.
-there are no ``gaps" in the chain.
-\end{defn}
-Observe that in a totally ordered set, there is only one maximal chain, namely the entire set.
-
-\begin{defn}
-A theoretical domain $\tdomain$ is said to have the \textbf{maximal chains property} (MCP) if given any chain $S\subseteq\tdomain$ totally ordered by narrowness ($\narrower$), the following holds.
-There exists a maximum chain $S'\supseteq S$, and a maximal chain $C$ in the set of equigranularity equivalence classes, such that for all statements $s \in S'$, there exists an equivalence class $\overline{c}\in C$ with $s\in \overline{c}$.
-\end{defn}
-
-\begin{prop}
-Suppose there is a (partial) pre-ordering on the countably infinite proper subsets of $\mathbb{N}$ which satisfies offset monotonicity.
-Then there exists a uniform theoretical domain which lacks the maximal chains property.
-\end{prop}
-\begin{proof}
-Consider the following example.
-Let the set of possibilities be given by $\mathbb{N}$ and let the theoretical domain be the powerset $\mathcal{P}(\mathbb{N})$.
-Let fineness be defined as follows.
-All finite subsets are ordered by their cardinality (with $\impossibility = \emptyset$).
-Put any ordering on the countably infinite strict subsets of $\mathbb{N}$, and impose that they are strictly coarser than all finite sets and strictly finer than the entire set $\mathbb{N} = \certainty$.
-This is an example of a uniform theoretical domain, where all possibilities (natural numbers) are equigranular (singleton sets of cardinality 1), and the ordering of equigranular equivalence classes is total.
-
-Consider the following chain $C$ in $\tdomain$: $c_0 = \emptyset$, $c_n = \{1,\ldots, n\}$ for $n<\infty$ and write $c_{\infty} = \mathbb{N}$.
-Then $c_m \snarrower c_n$ for all $m<n$.
-We claim that this is a maximal chain (see below for an explanation).
-
-Now, notice that there are no infinite proper subsets of $\mathbb{N}$ which are in the chain $C$.
-This represents an equigranularity class which is not represented in the maximal chain $C$, which is a failure of the maximal chains condition.
-
-To see why $C$ is maximal, let $c\in\mathcal{P}(\mathbb{N})$ but that $c\neq c_{\lambda}$ for $\lambda\in\mathbb{N}\cup\{0,\infty\}$.
-Then there exists some $l\in\mathbb{N}$ with $l+1\in c$ but $l\notin c$.
-Thus $c$ is not broader than $c_p$ for any $p\geq l$, and $c$ is not narrower than $c_q$ for any $q\leq l+1$.
-It follows that for $c$ to fit in the chain, it must be higher up than $c_{l+1}$ but lower down that $c_l$, which is not possible.
-\end{proof}
-
-\begin{prop}\label{chainimpliesuniform}
-Let $\tdomain$ be a theoretical domain with the maximal chains property. Then the following are equivalent:
-\begin{enumerate}
-    \item the possibilities are comparable
-    \item the possibilities are equigranular
-    \item all statements compatible with only finitely many possibilities are comparable to each other.
-\end{enumerate}
-\end{prop}
-\begin{proof}
-
-Suppose all possibilities are comparable.
-Let $p,q$ be possibilities.
-Consider the two chains $\{\impossibility, p,\certainty\}$ and $\{\impossibility, q, \certainty\}$.
-By the MCP, these may be extended to maximal chains in narrowness, with corresponding maximal chains in fineness equivalence classes.
-Then since possibilities are successors of the impossibility, the fineness equivalence classes represent by $p$ and $q$ must both be an immediate successor of the equivalence class representing the impossibility.
-Now, since all possibilities are comparable, and the fineness class of $p$ is an immediate successor to that of the impossibility, we must have $p\finer q$.
-Similarly, we must have $q\finer p$.
-It follows that $p\eqgran q$.
-
-Assume all possibilities are equigranular.
-We show that any two statements compatible with only finitely many possibilities are comparable.
-In fact, we will prove that the equigranularity equivalence classes for finite disjunctions of possibilities are given by the size of the disjunction.
-We proceed by induction on $N$, the larger number of compatible possibilities for the two statements.
-The case of $N=1$ follows from equigranularity of possibilities.
-Now suppose the claim holds for some $N$.
-Let $\stmt_1 = p_1\OR\cdots\OR p_{N+1}$ and $\stmt_2 = q_1\OR\cdots\OR q_m$ be statements compatible with only finitely many possibilities $p_i,i=1,\ldots,N+1$ and $q_j,j=1,\ldots,m$ respectively, where $m\leq N+1$.
-
-Consider first the case where $m\leq N$.
-Then $\stmt_2 \finer p_1\OR\cdots\OR p_N \sfiner p_1\OR\cdots\OR p_N\OR p_{N+1} = \stmt_1$, so they are comparable, and in particular all statements compatible with fewer than $N+1$ possibilities are strictly finer than $\stmt_1$.
-Next, suppose $m=N+1$.
-Write $\stmt_1' = p_1\OR\cdots\OR p_N$ and $\stmt_2' = q_1\OR\cdots\OR q_N$.
-We have already that all disjunctions of $N$ possibilties are equigranular, so $\stmt_1'\eqgran\stmt_2'$.
-Possibly by rearranging the order of the possibilties in the disjunction, we may assume $q_{N+1}\ncomp\stmt_1'$ and $q_{N+1}\ncomp\stmt_2'$.
-By Proposition \ref{fineness_properties}, we have $\stmt_1\OR q_{N+1}\eqgran\stmt_2$.
-Similarly, since $p_{N+1}\eqgran q_{N+1}$, we have $\stmt_1\eqgran\stmt_1'\OR q_{N+1}$.
-We conclude that $\stmt_1\eqgran\stmt_2$.
-This completes the proof that $2\Rightarrow3$.
-
-The claim $3\Rightarrow1$ is immediate since 1 is a special case of 3. 
-\end{proof}
-
-\begin{remark}
-The above observations and propositions show that MCP and comparability of possibilities imply equigranular possibilities, but equigrangular possibilities and totality do \emph{not} imply MCP.
-
-Further, MCP alone implies neither totality nor equigranular possibilities.
-Indeed, consider a theoretical domain with only two possibilities and a fineness ordering which has two unequal successors to the fineness of $\impossibility$.
-This satisfies the MCP but fails the other conditions.
-\end{remark}
-
-\begin{prop}\label{uniformconsequence}
+\begin{prop}
     Let $\tdomain$ be uniform.
-Let $\stmt, \stmt_1 \in \tdomain$ such that $\stmt \sfiner \stmt_1$.
+Let $\stmt, \stmt_1 \in \tdomain$ such that $\stmt \snarrower \stmt_1$.
 Then there exist $\stmt_2 \in \tdomain$ such that $\stmt_2 \narrower \stmt_1$ and $\stmt \eqgran \stmt_2$.
 \end{prop}
-
-The start of a proof....
-Let $S = \{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$. We will find a maximal element with respect to fineness (OR narrowness??).
-
-The difficulty here is that for Zorn we need to find an upper bound of an arbitrary chain which also lies in $S$.
-But such an upper bound is entirely unclear on how to obtain without using the proposition to begin with.
-Uniformity is difficult to make much use of for infinite domains (beyond that $S$ is nonempty).
-
-This set $S$ is nonempty, since any possibility narrower than $\stmt_1$ must also be finer than $\stmt$ by uniformity, and the statement $\stmt_1$ must not be the impossibility since it is strictly coarser than the statement $\stmt$.
-
-Now, given any chain (with respect to narrowness), consider its disjunction. But again this may be coarser than $\stmt$!
-
-Note: problem with chains for fineness.
-Taking the disjunction of a chain with respect to fineness can yield something much coarser (i.e.
-outside of $\{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$) if they do not overlap.
-
-Idea: given chains in fineness, can we construct new chains which are, at each level, equigranular, but now they are also chains in narrowness?
-
-Lemma: If $C$ is a chain in fineness and $\stmt$ is such that $\stmt[t]\finer\stmt$ for all $\stmt[t]\in C$, then there exists a set $N$ and a bijection $f:C\to N$ such that $N$ is a chain in narrowness, $\OR N \narrower \stmt$, and for all $\stmt[t]\in C$, $\stmt[t]\eqgran f(\stmt[t])$.
-Use this to move chains in fineness over to chains in narrowness and then can find maximal elements in the set $S$ above ``up to equigranular equivalence classes".
-
-Problem: this is actually much harder. Prop 6 is the case of this Lemma for a chain of length 2 and the top of the chain being equal to $\stmt$.
 
@@
 Having seen that, we can define infinitesimal in the following way.
-
-\begin{defn}\label{infinitesimals}
-Let $\stmt[u],\stmt[s] \in \tdomain$.
-Then $\stmt$ is \textbf{infinitesimal} with respect to $\stmt[u]$ if for any natural number $n$ we can find a set of $n$ statements $\{ \stmt_i \}_{i=1}^n$ such that:
+Let $\stmt[u],\stmt[s] \in \tdomain$ such that $\stmt \narrower \stmt[u]$.
+Then $\stmt$ is infinitesimal with respect to $\stmt[u]$ if for any natural number $n$ we can find a set of $n$ statements $\{ \stmt_i \}_{i=1}^n$ such that:
 \begin{enumerate}
@@
 In words, $\stmt$ is infinitesimal with respect to $\stmt[u]$ if for any number, there is a disjunction of that many broader mutually incompatible statements $\stmt_i$ is narrower than $\stmt[u]$.
-\end{defn}
-
-\begin{coro}
-    The impossibility $\impossibility$ is an infinitesimal with respect to any $\stmt[u] \in \tdomain$.
-\end{coro}
-
-\begin{proof}
-    Let $\stmt[u] \in \tdomain$ and consider $\{ \stmt_i \}_{i=1}^n$ where $\stmt_i \equiv \impossibility$ for all $i$.
-Then $\impossibility \finer \impossibility \equiv \stmt_i$ for all $i$.
-Since $\impossibility \ncomp \impossibility$, $\stmt_i \ncomp \stmt_j$ for all $i$ and $j$.
-Lastly, $\bigOR\limits_{i=1}^n \stmt_i \equiv \impossibility \finer \stmt[u]$.
-This means that $\{ \stmt_i \}_{i=1}^n$ realizes infinitesimality for $\impossibility$, which is an infinitesimal of $\stmt[u]$.
-\end{proof}
-
-\begin{prop}\label{infinitesimal_conjunction}
-Let $\stmt[v] \in \tdomain$ be an infinitesimal with respect to $\stmt[u] \in \tdomain$. Then any $\stmt \finer \stmt[u]$ is also an infinitesimal with respect to $\stmt[u]$.
-\end{prop}
-\begin{proof}
- Since $\stmt[v]$ is an infinitesimal with respect to $\stmt[u]$, for any $n$ we can find a set of statements $\{ \stmt_i \}_{i=1}^n$ that realizes its infinitesimality.
-We have $\stmt \finer \stmt[v] \finer \stmt_i$ for all $i$.
-The other two properties do not depend on $\stmt[v]$.
-Therefore $\{ \stmt_i \}_{i=1}^n$ realizes infinitesimality for $\stmt$ as well so $\stmt$ is an infinitesimal with respect to $\stmt[u]$.
-\end{proof}
-
-\begin{coro}
-The conjunction of two infinitesimals with respect to $\stmt[u]$ is infinitesimal with respect to $\stmt[u]$.
-\end{coro}
-\begin{proof}
-Let $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ be infinitesimal with respect to $\stmt[u]$.
-Since $\stmt[v]_1 \AND \stmt[v]_2 \narrower \stmt[v]_1$, $\stmt[v]_1 \AND \stmt[v]_2 \finer \stmt[v]_1$ and therefore $\stmt[v]_1 \AND \stmt[v]_2$ by \ref{infinitesimal_conjunction}.
-\end{proof}
-
-\begin{prop}
-The disjunction of two infinitesimals with respect to $\stmt[u]$ that are comparable\footnote{Is comparability required?
-Can it be proved one way or another?} to each other is infinitesimal with respect to $\stmt[u]$.
-\end{prop}
-\begin{proof}
-Let $\stmt[v],\stmt[w]$ be infinitesimal with respect to $\stmt[u]$.
-Without loss of generality, suppose $\stmt[v] \finer \stmt[w]$.
+Notice that if a set of size $n$ exists, then for all $k$ with $k<n$ a set must also exist, so in the definition it would be equivalent to say, for arbitrarily large $n$, such a set $\{ \stmt_i \}_{i=1}^n$ exists.
+
+If $\stmt[v]$ is not infinitsimal with respect to $\stmt[u]$, then there exists a natural number $n$ such that any collection of $n$ statements must have at least one of the following: one (or more) of them is strictly finer than $\stmt[v]$; some pair of them is compatible; or their disjunction is strictly coarser than $\stmt[u]$.
+
+This will work if all possibilities are equigranular.
+It will not work in general.
+For example, we could have 3 possibilities $\{a,b,c\}$ with $P(a) = 0$ and $P(b)=P(c)=1/2$.
+But in that case the ordering would not be enough to recover the measure.
+
+\begin{prop}
+Let $\stmt[u]\in\tdomain$. Finite disjunctions of statements infinitesimal with respect to $\stmt[u]$ are still infinitesimal with respect to $\stmt[u]$.
+\end{prop}
+\begin{proof}
+We show it for the case of a disjunction of two infinitesimals; other finite disjunctions will follow by induction.
+Let $\stmt,\stmt[t]$ be infinitesimal with respect to $\stmt[u]$.
+Consider first the case that $\stmt\finer\stmt[t]$ (this is identical to the case that $\stmt[t]\finer\stmt$).
+
 Let $n>0$.
-Let $\{\stmt[w]_i\}_{i=1}^{2n}$ be a collection realizing infinitesimality for $2n$ for $\stmt[w]$.
-Define $\stmt_j = \stmt[w]_j\OR\stmt[w]_{j+n}$ for $j=1,\ldots,n$.
-Notice that the $\stmt_j$'s are mutually incompatible, and their disjunction must be finer than $\stmt[u]$.
-Further, since $\stmt[v]\finer\stmt[w]$ we have that $\stmt[v]\finer\stmt[w]_i$ for all $i$.
+Let $\{\stmt[t]_i\}_{i=1}^{2n}$ be a collection realizing infinitesimality for $2n$ for $\stmt[t]$.
+Define $\stmt[v]_j = \stmt[t]_j\OR\stmt[t]_{j+n}$ for $j=1,\ldots,n$.
+Notice that the $\stmt[v]_j$'s are mutually incompatible, and their disjunction must be finer than $\stmt[u]$.
+Further, since $\stmt\finer\stmt[t]$ we have that $\stmt\finer\stmt[t]_i$ for all $i$.
 It follows that:
 $$
-\stmt[v]\OR\stmt[w] \finer \stmt[w]_j\OR\stmt[w]_{j+n} = \stmt_j
+\stmt\OR\stmt[t] \finer \stmt[t]_j\OR\stmt[t]_{j+n} = \stmt[v]_j
 $$
 for all $j=1,\ldots,n$.
-Thus the collection $\{\stmt_j\}_{j=1}^n$ realizes $\stmt[v]\OR\stmt[w]$ being infinitesimal with respect to $\stmt[u]$ for $n$.
-We conclude that $\stmt[v]\OR\stmt[w]$ is infinitesimal with respect to $\stmt[u]$.
+Thus the collection $\{\stmt[v]_j\}_{j=1}^n$ realizes $\stmt\OR\stmt[t]$ being infinitesimal with respect to $\stmt[u]$ for $n$.
+We conclude that $\stmt\OR\stmt[t]$ is infinitesimal with respect to $\stmt[u]$.
+
+Consider next the case that $\stmt$ and $\stmt[t]$ are not comparable with respect to fineness.
+{\bf By totality, this case doesn't exist.}\footnote{I'm not quite sure how to show this when they're not comparable.
+It may be true and not too hard, I just did not see how to do it yet.}
+
 \end{proof}
@@
 
-\subsubsection{Equivalence at u scale}
-
-\def\equivu{\cong_{\stmt[u]}}
-
-\begin{defn}
-Let $\stmt_1, \stmt_2 \in \tdomain$.
-We say $\stmt_1$ is \textbf{equivalent to $\stmt_2$ at $\stmt[u]$ scale}, noted $\stmt_1 \equivu \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$.
-\end{defn}
-
-\begin{prop}
-Equivalence at $\stmt[u]$ scale is an equivalence class.
-\end{prop}
-\begin{proof}
-For reflexivity, let $\stmt \in \tdomain$.
-We have $\stmt \OR \impossibility \equiv \stmt \OR \impossibility$.
-Since $\impossibility$ is an infinitesimal for all $\stmt[u]$ (TODO: add proposition), $\stmt \equivu \stmt$.
-For symmetry, note that the definition itself is symmetrical.
-For transitivity, let $\stmt_1 \equivu \stmt_2 \equivu \stmt_3$.
-Then we have $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$ and $\stmt_2 \OR \stmt[w]_2 \equiv \stmt_3 \OR \stmt[w]_3$.
-\end{proof}
-
 \subsubsection{Greatness in $\stmt[u]$}
@@
 \end{prop}
-\begin{proof}[Using Prop \ref{uniformconsequence}]
-If both of the $\finer$'s are actually equigranular, there is nothing to show.
-Next consider the case that $\stmt_1\sfiner\stmt\finer\stmt_1\OR\stmt[v]_1$.
-Then by Proposition \ref{uniformconsequence} there exists $\stmt_2$ such that $\stmt_2\narrower\stmt$ and $\stmt_2\eqgran\stmt_1$.
-But how do we put this together??
-What about the even simpler version of the lemma with $\narrower$ instead of $\finer$?
-Just use $\stmt\OR (\stmt_1\OR\stmt[v]_1 \AND \NOT\stmt)$.
-\end{proof}
 
@@
 
-\section{Possible example}
-
-Let $\mathbb{N}$ be the set of natural numbers.
-We define the sets $D_{i,k} = \{ x \in \mathbb{N} \, | \, x \mod i = k \}$ with $k < i$.
-Therefore $D_{1,0} = \mathbb{N}$, $D_{2,0} = \{ 2, 4, 6, ...
-\}$, $D_{2,1} = \{ 1, 3, 5, ...
-\}$ and so on.
-Additionally, we say $D_{i,j} \finer D_{k,l}$ for all $k \leq i$ and for all $j$ and $l$.
-As needed, we may overload the notation by letting $D_{i,j} =  D_{i, j \mod i}$ where $j\mod i$ is the smallest positive residue of $j$ under $i$.
-
-Let $D_{i,j}$ and $D_{k,l}$ be two of these subsets.
-Consider the intersection $D_{i,j}\cap D_{k,l}$ and suppose it is nonempty; let $n$ be in this intersection.
-Then $D_{i,j} = D_{i,n}$ and $D_{k,l} = D_{k,n}$, so the intersection is given by $D_{i,n}\cap D_{j,n} = \{x\in \mathbb{N} \, | \, x \mod \gcd(i,j) = n\} = D_{\gcd(i,j),n}$.
-
-Note also that $D_{i,j}^C = \cup_{k=1,k\neq j}^{i-1}D_{i,k}$.
-
-It follows that any finite combination of unions, intersections, and complements of sets of the form $D_{i,j}$ will be empty, or countably infinite.
-
-Next, let $n\in\mathbb{N}$.
-Consider the union $S_n = \cup_{i=n+1}^{\infty}D_{i,0} = \{n+1,n+2,\ldots\}$, which may be generated by a countable union.
-Thus $S_n^C = \{1,2,\ldots,n\}$.
-By combining complements and intersections, we can obtain $\{n\} = S_{n-1}\cap S_n^C$.
-We can then generate arbitrary subsets of $\mathbb{N}$ using countable operations of union, intersection, and complement on the sets $D_{i,j}$.
-
-Claim: all finite operations of union, intersection, and complement of $D_{i,j}$'s yield sets that may be written as finite unions of $D_{i,j}$'s.
-Countable operations yield arbitrary subsets of $\mathbb{N}$.
-
-Possible source to read from: http://www.stat.cmu.edu/tr/tr814/tr814.pdf
-Note in abstract, gives $1/m$ probability (measure) to $D_{m,j}$ for each $j$. 
 \section{Other notes}

### 2021-02-22 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Edits on the action principle

@@
 Yet, as Section~\ref{dependence} explained, the density is always coordinate-dependent, even when coordinate-invariant.
-In other words, there is no coordinate-independent (or coordinate-free definition) of the invariant phase space density.
+In other words, there is no coordinate-independent (or coordinate-free) definition of the invariant phase space density.
 Affording the density fundamental significance therefore places us at odds with those who take coordinate-independence as a necessary condition for fundamentality.
@@
 [XXX: check consistency with previous claim] By being gauge-variant, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can remain physically significant but not fundamental).
+[XXX: still, conjugate momentum is less physically significant because is not well defined experimentally]
 
@@
 
+The setting is the extended phase space $(q, p, t)$, which is the standard phase space extended by the time variable.
 Knowing that the system is deterministic and reversible, it follows that the density of states is invariant.
-This means that possible trajectories of the system are never pushed toward each other or pulled apart.
-In other words, the flux of trajectories through any closed surface is always zero.
-Formally, the displacement of the states in time is divergenceless: $\nabla \cdot \vector{S} = 0 $.
-Here, $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ defines a vector field in the extended phase space $(q, p , t ) $.
-$\vector{S}$ characterizes the displacement of states in time.
-In particular, it is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and affine parameter of the trajectory.
+This means that possible trajectories of the system in the extended phase space are never pushed toward each other or pulled apart over time.
+In other words, the flux of trajectories through any closed surface in the extended phase space is always zero.
+Formally, the vector field $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ represents the displacement of states along the trajectories.
+This field is divergenceless, $\nabla \cdot \vector{S} = 0$, because, given a region $U$ as many states will enter the region as exit.
+In particular, note that the displacement is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and the parameter of the trajectory.
+[XXX:would avoid using affine parameter...
+it may not be affine]
 
@@
 
-Treating time $t$ as both a coordinate and an affine parameter (parameterizing the trajectory), the time component of the displacement, $S_t = dt/ dt = 1$.\footnote{Considering $dt/dt$, the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
+Treating time $t$ as both a coordinate and the parameter of the trajectory, the time component of the displacement, $S_t = dt/ dt = 1$.\footnote{Considering $dt/dt$, the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
 We could instead use a parameter $\tau$, and the derivation would remain valid.} This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $.
@@
 Hence, we write $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
-By matching components of $\vector{S}$ and $-curl(\theta)$, we show that H is the Hamiltonian for the system.
+By matching components of $\vector{S}$ and $-curl(\theta)$, we show that $H$ is the Hamiltonian for the system.
 First, by definition, $S_q \equiv \frac{dq}{dt} $ and $S_p \equiv  \frac{dp}{dt} $.
@@
 We notice that the integrand simply equals the inner product of $\vector{\theta}$ with $\vector{S} $, entailing that the Lagrangian $L =\vector{\theta} \cdot \vector{S} $.
-Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$.
-Changing variables, we take the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.\footnote{Note that $\vector{S}  dt = d\vector{\gamma}$ only on the actual trajectory in phase space, but not on other possible trajectories.} To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
+Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$ where $\gamma$ is an actual trajectory followed by the system.
+Note that $\vector{S}  dt = (\frac{d q }{d t } dt, \frac{d p }{d t } dt, \frac{d t }{d t }dt ) = (dq, dp, dt) = d\vector{\gamma}$ represents the infinitesimal displacement along the trajectory.
+Therefore the action for those trajectories can be rewritten as the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.
+To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
 % Note on the generality of the argument: in cases where motion equivalence is not satisfied, we can still write the integrand in terms of H(q, p, t), although it would not be an integral over the Lagrangian.
@@
 By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.
-For this ``system," all paths have the same action, so all paths are possible.
+[XXX: would remove mention of Hamiltonian] For this ``system," all paths have the same action, so all paths are possible.
 This gives an initial probability distribution over the states, but it does not specify how this initial distribution will evolve into a future distribution of states.
 In other words, the Lagrangian alone does not supply a probability distribution on the possible evolutions.
-In contrast, the Hamiltonian still defines a unique solution (i.e.
+[XXX: I would say, it does not tell us how an initial probability would evolve in time] In contrast, the Hamiltonian still defines a unique solution (i.e.
 a unique evolution in phase space) even when outside the hyper-regular domain (i.e.

### 2021-02-22 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Adding changes from meeting

@@
 We will call a set of quantities $\xi^a$ \textit{state variables} provided that they fully identify a state, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates'.
-However, we will have multiple manifolds and, to avoid confusion, we will call `state variables' the coordinates of the state space.} We define a density $\rho^{\ast}_c$: $\mathbb{R}^n \to \mathbb{R}$ as the pullback of the distribution $\rho_c$ by the inverse of a coordinate map:  $\rho^{\ast}_c \equiv (\rho_c \circ \xi^{-1})$.
+However, we will have multiple manifolds and, to avoid confusion, we will call `state variables' the coordinates of the state space.} We define a density $\rho_{c, \xi}$: $\mathbb{R}^n \to \mathbb{R}$ as the pullback of the distribution $\rho_c$ by the inverse of a coordinate map:  $\rho_{c,\xi} \equiv (\rho_c \circ \xi^{-1})$.
 %To represent the distribution over state variables, we define coordinate charts $(U, \xi)$ over $\mathcal{S}$.
@@
 
-The distribution $\rho_c$ is a function of the particle states $s \in \mathcal{S}$, while the density $\rho^{\ast}_c$ is a function of the state variables $\xi^a \in \mathbb{R}^n$.
-A coordinate map $\xi$ lets us change variables from $s$ to $\xi^a$.
-Consequently, we require that $\mu(U) = \int_U \rho_c(s) d\mathcal{S} = \int_{\xi(U)} \rho^{\ast}_c (\xi^a) d \xi^a$.\footnote{Indeed, using the Lebesgue measure $\lambda$ on $\mathbb{R}^n$---and assuming that $\xi^{-1}$: $\mathbb{R}^n \to \mathcal{S}$ is measurable---we can define the measure $d\mathcal{S}$ by the pushforward measure $\xi^{-1}_{\ast}(\lambda)(U) \equiv \lambda(\xi(U))$.} Since $\rho_c$: $\mathcal{S} \to \mathbb{R}$ is a function of the particle state $s$, its value is independent of the choice of state variables $\xi^a$.
-Hence, $\rho_c$ is a scalar function and consequently coordinate independent.
+The distribution $\rho_c$ is a function of the particle states $s \in \mathcal{S}$, while the density $\rho_{c, \xi}$ is a function of the state variables $\xi^a \in \mathbb{R}^n$.
+A coordinate map $\xi$ lets us express the state $s$ as a function of the coordinates $\xi^a$.
+Consequently, we require that $\mu(U) = \int_U \rho_c(s) d\mathcal{S} = \int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a$.\footnote{Indeed, using the Lebesgue measure $\lambda$ on $\mathbb{R}^n$---and assuming that $\xi^{-1}$: $\mathbb{R}^n \to \mathcal{S}$ is measurable---we can define the measure $d\mathcal{S}$ by the pushforward measure $\xi^{-1}_{\ast}(\lambda)(U) \equiv \lambda(\xi(U))$.} Since $\rho_c$: $\mathcal{S} \to \mathbb{R}$ is a function of the particle state $s$, its value is independent of the choice of state variables $\xi^a$.
+Hence, $\rho_{c, \xi}$ will transform as a scalar under change of variables.
 Thus, $\int_U \rho_c(s) d\mathcal{S}$ is invariant under coordinate changes $\xi$ to $\hat{\xi}$.
@@
 In particular, note that the displacement is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and the parameter of the trajectory.
-[XXX:would avoid using affine parameter...
-it may not be affine]
 
@@
 Nonetheless, reformulating Hamiltonian mechanics within the extended phase space $(q_i, p_i, t, E)$ provides one way to rectify these deficiencies.
-At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and an affine parameter within the usual Hamiltonian framework.
+At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and a time-like parameter within the usual Hamiltonian framework.
 Yet, differential geometry demands that we separate these two roles.
 We thereby add a temporal degree of freedom $(t, E) $, expanding the phase space degrees of freedom by one (recall that we previously had one degree of freedom for every canonical coordinate pair $q_i, p_i$).
-Next, we define the phase space evolution over an affine parameter $s $.
+Next, we define the phase space evolution over an evolution parameter $s $.
 This yields two more equations of motion: $dt/ds = -\partial \mathcal{H}/ \partial E$ and $dE/ds = -\partial \mathcal{H}/ \partial t$, where $\mathcal{H}(q_i, p_i, t, E)$ is the extended Hamiltonian.
-In this framework, there is a \textit{Hamiltonian constraint} $\mathcal{H}=0=H(q_i, p_i, t)-E$ that functions as both a global constraint---fixing the energy as a function of $q, p, t $---and as a generator of the affine parameter $s $ [XXX to do: provide a reference for extened phase space formulation.
+The extended Hamiltonian not only acts as a generator of the evolution parameter $s$, but it also provides us with the \textit{Hamiltonian constraint} $\mathcal{H}=0$ [XXX to do: provide a reference for extened phase space formulation.
 And check this discussion of the `Hamiltonian constraint'].
+The Hamiltonian constraint provides the relationship between the energy and the other state variables.
 Crucially, the extended Hamiltonian---unlike the ordinary Hamiltonian---is coordinate-invariant, even with respect to time-dependent coordinate transformations.

### 2021-02-23 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Incorporated changes from 2-22 meeting

@@
 
-In the continuous case, when $\mathcal{S}$ is a manifold, we define $\mu(U) = \int_U \rho_c(s) d\mathcal{S} $.
-We will call a set of quantities $\xi^a$ \textit{state variables} provided that they fully identify a state, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates'.
-However, we will have multiple manifolds and, to avoid confusion, we will call `state variables' the coordinates of the state space.} We define a density $\rho_{c, \xi}$: $\mathbb{R}^n \to \mathbb{R}$ as the pullback of the distribution $\rho_c$ by the inverse of a coordinate map:  $\rho_{c,\xi} \equiv (\rho_c \circ \xi^{-1})$.
-%To represent the distribution over state variables, we define coordinate charts $(U, \xi)$ over $\mathcal{S}$.
-Likewise, the distribution $\rho_c$ is the pullback of $\rho^{\ast}_c$ by a coordinate map: $\rho_c (s) = (\xi^{\ast} \rho^{\ast}_c)(s) = \rho^{\ast}_c (\xi (s))$.
-
-The distribution $\rho_c$ is a function of the particle states $s \in \mathcal{S}$, while the density $\rho_{c, \xi}$ is a function of the state variables $\xi^a \in \mathbb{R}^n$.
-A coordinate map $\xi$ lets us express the state $s$ as a function of the coordinates $\xi^a$.
-Consequently, we require that $\mu(U) = \int_U \rho_c(s) d\mathcal{S} = \int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a$.\footnote{Indeed, using the Lebesgue measure $\lambda$ on $\mathbb{R}^n$---and assuming that $\xi^{-1}$: $\mathbb{R}^n \to \mathcal{S}$ is measurable---we can define the measure $d\mathcal{S}$ by the pushforward measure $\xi^{-1}_{\ast}(\lambda)(U) \equiv \lambda(\xi(U))$.} Since $\rho_c$: $\mathcal{S} \to \mathbb{R}$ is a function of the particle state $s$, its value is independent of the choice of state variables $\xi^a$.
-Hence, $\rho_{c, \xi}$ will transform as a scalar under change of variables.
-Thus, $\int_U \rho_c(s) d\mathcal{S}$ is invariant under coordinate changes $\xi$ to $\hat{\xi}$.
-As a function of the state variables $\xi^a$, $\rho^{\ast}_c$ transforms according to the Jacobian for this change of variables: $\rho^{\ast}_c (\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho^{\ast}_c (\hat{\xi}^b)$.
-Since $\mu(U) = \int_{\xi(U)} \rho^{\ast}_c (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho^{\ast}_c (\hat{\xi}^b) d \hat{\xi}^b$ remains invariant under this coordinate change, the Jacobian determinant must equal one.
-By dimensional homogeneity, it must also be dimensionless.\footnote{Formally, we can express this argument as a chain of equalities: $\rho^{\ast}_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho^{\ast}_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho^{\ast}_c(\hat\xi^b)$, where $\rho^{\ast}_c(\xi^a)$ and $\rho^{\ast}_c(\hat{\xi}^b)$ are the same density expressed over different state variables and $s(\xi^a) \equiv \xi^{-1}(\xi^a)$ and $s(\hat{\xi}^b)\equiv \hat{\xi}^{-1}(\hat{\xi}^b)$ represent the same particle state as expressed by these different variables.
-Note that $\rho_c(s(\hat{\xi}^b))$ shows how the distribution depends on the state variables only through their labeling of a particle state $s$.}
-%$\rho^{\ast}_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho^{\ast}_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho^{\ast}_c(\hat\xi^b)$
-
-We will now show that in order for the distribution $\rho_c$ to satisfy these two constraints, the classical phase space must be at least a symplectic manifold.
-In other words, having symplectic manifold structure is a necessary condition for the distribution over particle states to transform as a scalar function and for its pullback $\rho^{\ast}_c := \rho_c \circ \xi^{-1}$ to transform as a density.
-
+In the continuous case, $\mathcal{S}$ is a manifold, and we define $\mu(U) = \int_U \rho_c(s) d\mathcal{S} $.
+We will call a set of quantities $\xi^a$ \textit{state variables} provided that they fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates' and $\xi$ is a coordinate map.
+We will have multiple manifolds, so to avoid confusion, we will call the coordinates of the state space `state variables'.
+As usual with coordinate charts, $\xi$ is properly defined on some $U \subseteq \mathcal{S}$, and $\xi^{-1}$ is defined on $\xi(U) \subseteq \mathbb{R}^n$.
+We will neglect these subtleties in what follows.} In other words, $\xi$ represents a state $s \in \mathcal{S}$ by a set of state variables $\xi^a \in \mathbb{R}^n$, such that $\xi (s) = \xi^a$.
+To represent the distribution $\rho_c$ in terms of state variables, we define a density $\rho_{c, \xi}$: $\mathbb{R}^n \to \mathbb{R}$ by $\rho_{c, \xi} \equiv (\rho_c \circ \xi^{-1})$.
+This density plays a crucial role throughout our argument.
+% A state variable change $\hat{\xi}$ represents the same state $s$ by a different set of state variables $\hat{\xi}^b$.
+
+Since $\rho_c$: $\mathcal{S} \to \mathbb{R}$ is a function of the particle state $s$, its value is independent of the choice of state variables $\xi^a$ (i.e.
+the choice of state variable map $\xi$).
+Hence, the density $\rho_{c, \xi}$ transforms as a scalar under state variable changes.\footnote{For two state variable maps $\xi$ and $\hat{\xi}$, we have $\rho_c(s) = (\rho_c \circ \xi^{-1} \circ \xi)(s) = (\rho_c \circ \hat{\xi}^{-1} \circ \hat{\xi})(s)$, which entails that $\rho_c(s) = \rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
+Here, $\xi(s) \equiv \xi^a$ and $\hat{\xi}(s) \equiv \hat{\xi}^b$.} Likewise, the integral $\int_U \rho_c(s) d\mathcal{S} $ is independent of the choice of state variables, so $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.
+By the change of variables formula, we also know that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a= \int_{\hat{\xi}(U)} \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.
+Here, $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ is the determinant of the Jacobian for this change of state variables.
+Combining these two equations, we see that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b = \int_{\hat{\xi}(U)} \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_{c, \hat{\xi}} (\hat{\xi}^b)d \hat{\xi}^b$, entailing that the Jacobian determinant must equal one.
+By dimensional homogeneity, it must also be dimensionless.
+We will now show that in order for the density $\rho_{c, \xi}$ to have these transformation properties, the classical phase space must be at least a symplectic manifold.
+%footnote{Indeed, using the Lebesgue measure $\lambda$ on $\mathbb{R}^n$---and assuming that $\xi^{-1}$: $\mathbb{R}^n \to \mathcal{S}$ is measurable---we can define the measure $d\mathcal{S}$ by the pushforward measure $\xi^{-1}_{\ast}(\lambda)(U) \equiv \lambda(\xi(U))$.}
 %In other words, while we can change the units used by the state variables $\xi^a$ to identify the particle states, the units for $\rho^{\ast}_c$ must remain the same so that the density does not change its numeric value
@@
 This agrees with the practice in statistical mechanics.
-The density $\rho_c$ [XXX: distribution or density?], when normalized, is given in terms of the inverse of the units for $\omega$.
+The density $\rho_{c, \xi}$, when normalized, is given in terms of the inverse of the units for $\omega$.
 In accordance with the standard conventions in quantum mechanics and statistical mechanics, the $k_i$ are given in terms of inverse units of the respective $q^i$, such that $dq^i \wedge dk_i$ is a pure number.
@@
 In the remainder of this essay, we provide further reasons for privileging Hamiltonian mechanics.
-The next section describes how Hamiltonian mechanics provides the natural setting for the invariant density, describing the matter content of the theory [XXX: matter content of the theory seems like an odd phrase].
+The next section describes how Hamiltonian mechanics provides the natural setting for the invariant density, which characterizes how a classical object's parts are distributed.
 Section~\ref{fundamental} argues that the invariant phase space density is more fundamental than the state variables, since conjugate momentum is a gauge-dependent quantity.
@@
 
-[XXX To do: insert discussion of statistical mechanics and entropy here: This is why statistical mechanics works best on phase space (we have a uniform measure over q and p).
-Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
-Areas/volumes are measures of number of states.
+[XXX To do: insert discussion of statistical mechanics and entropy here: statistical mechanics works best on phase space because phase space gives us a uniform measure over q and p.
+Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more salient/perspicuous.
+Areas/volumes are measures of the number of states.
 They are coordinate invariant.
-Also the Shannon and Gibbs entropies are coordinate invariant.]
+Also the Shannon and Gibbs entropies are coordinate invariant.]\footnote{For further discussion of the importance of phase space for statistical mechanics and entropy, see \textcites{shpmp} [omit this for review?].}
 
@@
 Being necessary for fundamental significance does not, however, make coordinate-invariance necessary for \textit{non-fundamental} physical significance.
-[XXX: disagree] Some quantities and properties are coordinate-variant yet still physically significant.\footnote{Indeed, many physically significant quantities are gauge-dependent, such as conjugate momentum and the Lagrangian (see Sections~\ref{gauge_momentum} and~\ref{gauge}).
+Many representational choices are coordinate-variant yet still physically significant.\footnote{Indeed, many physically significant quantities are gauge-dependent, such as conjugate momentum and the Lagrangian (see Sections~\ref{gauge_momentum} and~\ref{gauge}).
 Insofar as gauge-dependency is analogous to being coordinate-variant, this supplies a further reason for denying that coordinate-invariance is a necessary condition for physical significance.} Our position here again places us at odds with others in the literature.
 For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
-However, some choices of unit-variables $q$ are not appropriately connected with physical measurement and hence are less physically meaningful [XXX to do: provide a clear example of this ---  not sure we can make this claim].
+However, some choices of coordinates are particularly well-suited to certain problems, such as the use of action--angle variables for integrable systems, or spherical coordinates for systems with spherical symmetry.
 Similarly, we saw in Section~\ref{dependence} that some coordinate choices in Lagrangian mechanics are physically significant because they enable the property of point-wise comparison.
@@
 This provides another instance of a physically significant but coordinate-dependent representational choice.
-Indeed, the next section shows that the conjugate momentum is
+Similarly, the next section shows that the conjugate momentum is gauge-dependent, but presumably still physically significant.
 
@@
 Insofar as coordinate-invariance is a plausible necessary condition for fundamental physical status, so is gauge-invariance.
-[XXX: check consistency with previous claim] By being gauge-variant, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can remain physically significant but not fundamental).
-[XXX: still, conjugate momentum is less physically significant because is not well defined experimentally]
+By being gauge-dependent, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can remain physically significant but not fundamental).
+%[XXX: still, conjugate momentum is less physically significant because is not well defined experimentally]
 
@@
 In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
+Unlike kinetic momentum, conjugate momentum is not an experimentally measurable quantity, because its specification involves a gauge choice.
 
@@
 Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition.
-We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument illustrating this generalization, see \textcites[236--37, 243--45]{Arnold}.}
+We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument, see \textcites[236--37, 243--45]{Arnold}.}
 %our derivation seems very similar to Arnold chapter 9, section 45:C: ``the principle of least action in phase space''. Note that we briefly discussed Arnold on 11-23-20 meeting, around 12:30. 
@@
 
-The setting is the extended phase space $(q, p, t)$, which is the standard phase space extended by the time variable.
+The derivation uses the extended phase space $(q, p, t)$, i.e.
+the standard phase space extended by the time variable.
 Knowing that the system is deterministic and reversible, it follows that the density of states is invariant.
@@
 Formally, the vector field $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ represents the displacement of states along the trajectories.
-This field is divergenceless, $\nabla \cdot \vector{S} = 0$, because, given a region $U$ as many states will enter the region as exit.
-In particular, note that the displacement is tangent to the trajectory $\gamma [q, p, t] (t)$, in extended phase space, where the variable $t$ functions as both a coordinate and the parameter of the trajectory.
+This field is divergenceless, $\nabla \cdot \vector{S} = 0$, because given a region $U$ as many states will enter the region as exit.
+In particular, note that the displacement is tangent to the trajectory $\gamma [q, p, t] (t)$ in extended phase space, where the variable $t$ functions as both a coordinate and the evolution parameter.
 
@@
 
-(XXX to do: Note how the geometry is better understood in extended phase space, requiring the Hamiltonian picture.
+%(XXX to do (might no longer be necessary given earlier and later remarks) Note how the geometry is better understood in extended phase space, requiring the Hamiltonian picture.
 Note that this applies even when $\theta \cdot S$ cannot be expressed as a function of $q$ and $\dot q$.
@@
 
-Nonetheless, reformulating Hamiltonian mechanics within the extended phase space $(q_i, p_i, t, E)$ provides one way to rectify these deficiencies.
-At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and a time-like parameter within the usual Hamiltonian framework.
+Nonetheless, reformulating Hamiltonian mechanics within the extended phase space $(q_i, p_i, t, E)$ provides one way to rectify these deficiencies.\footnote{For a detailed discussion of the extended phase space formalism, see \textcites[]{Struckmeier}.} At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and a time-like parameter within the usual Hamiltonian framework.
 Yet, differential geometry demands that we separate these two roles.
@@
 This yields two more equations of motion: $dt/ds = -\partial \mathcal{H}/ \partial E$ and $dE/ds = -\partial \mathcal{H}/ \partial t$, where $\mathcal{H}(q_i, p_i, t, E)$ is the extended Hamiltonian.
-The extended Hamiltonian not only acts as a generator of the evolution parameter $s$, but it also provides us with the \textit{Hamiltonian constraint} $\mathcal{H}=0$ [XXX to do: provide a reference for extened phase space formulation.
-And check this discussion of the `Hamiltonian constraint'].
+The extended Hamiltonian not only acts as a generator of the evolution parameter $s$, but it also provides us with the \textit{Hamiltonian constraint} $\mathcal{H}=0$.
 The Hamiltonian constraint provides the relationship between the energy and the other state variables.
 Crucially, the extended Hamiltonian---unlike the ordinary Hamiltonian---is coordinate-invariant, even with respect to time-dependent coordinate transformations.
-For example, if we define the affine parameter as the proper time, then the extended Hamiltonian is proportional to the mass (which is manifestly coordinate-invariant).
+For example, if we define the evolution parameter as the proper time, then the extended Hamiltonian is proportional to the mass (which is manifestly coordinate-invariant).
 This provides a clear physical meaning for the extended Hamiltonian.
@@
 For example, consider the case where the Lagrangian $L$ equals zero.
-This Lagrangian gives rise to a Hamiltonian that also equals zero.
-By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.
-[XXX: would remove mention of Hamiltonian] For this ``system," all paths have the same action, so all paths are possible.
-This gives an initial probability distribution over the states, but it does not specify how this initial distribution will evolve into a future distribution of states.
-In other words, the Lagrangian alone does not supply a probability distribution on the possible evolutions.
-[XXX: I would say, it does not tell us how an initial probability would evolve in time] In contrast, the Hamiltonian still defines a unique solution (i.e.
+For this ``system," all paths have the same action, so all paths are possible.
+Given an initial probability distribution over states, the Lagrangian fails to specify how this distribution will evolve in time.
+In contrast, a Hamiltonian still defines a unique solution (i.e.
 a unique evolution in phase space) even when outside the hyper-regular domain (i.e.
 when the Hessian is zero).\footnote{In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
-There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.} Indeed, Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}.
+There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.} For instance, a zero Hamiltonian describes a static system (by Hamilton's equations, the derivatives of the position and momenta are zero).
+Indeed, Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}.
 Curiel argues that this is an advantage of Lagrangian mechanics, but it generates challenges for understanding non-hyper-regular Lagrangians.\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.

### 2021-02-23 Overviews/Experimental verifiability/ExperimentalVerifiabilityOverview.tex

@@
 
-\usepackage{amsmath}
-\usepackage{amsfonts}
+\input{include/basicstyle}
+\input{include/logic}
+\input{include/theorems}
+
 \usepackage{graphicx}
-\usepackage{enumitem}
 \usepackage{hyperref}
@@
 
+The guiding principle is that \emph{science is universal, non-contradictory and evidence based.} This means that any scientific theory must specify a set of \textbf{statements} that are logically consistent, whose truth can be confirmed experimentally by everybody.
+This starting point is codified in the axioms that form the formal foundation of the theory.
+We start with the axioms of logic.
+
+The \emph{axiom of context} tells us that statements are organized into a \textbf{logical context} $\logCtx$, and a well defined truth value for each statement must exist.
+This recognizes that the logical consistency and meaning of each statement depends on definitions and relationships that depends on the context, on a set of statements.\footnote{For example, if one is performing particle identification, the mass of an electron is assumed to be given.
+If one is measuring the mass, the identify is assumed to be given, which the value of the mass is unknown.}
+
+The \emph{axiom of possibility} tell us that the statements in a context cannot be assigned arbitrary truth values, but only in a way that is consistent with the meaning defined by the context.\footnote{For example, \statement{the mass of the object is one kilogram} and \statement{the mass of the object is two kilograms} cannot be both true.} Therefore each context comes with a set $\pAss$ of possible assignments, where each possible assignment is a function $a : \logCtx \to \Bool$ that returns a value of truth for each statement.
+The truth is a possible assignment.
+
+The \emph{axiom of closure} tells us that we can always find a statements whose truth arbitrarily depends on the truth of other statements.
+
+From the logic axioms one can show that each logical context is a complete Boolean algebra; that the narroness operator $\narrower$ that tells us whether a statement is more specific than another\footnote{For example, \statement{the mass of the object is between 4 and 5 kilograms} is narrower than \statement{the mass of the object is between 0 and 100 kilograms}.} provides a partial order for the context; that we can define intuitive relationship between statements, such as compatibility or independence.
+
+Next, we introduce the axioms of verifiable logic.
+The \emph{axiom of verifiability} tells us that some statements are \textbf{verifiable}, meaning we have a test at our disposal that will terminate successfully in finite time if and only if the statement is true.
+Note that the test is not guaranteed to terminate, to give an answer, if the test is false.\footnote{For example, a test for \statement{there exists extra-terrestrial life} may terminate successfully if life is found, but it will not be conclusive if life is not found.} Therefore the negation of a verifiable statement is not necessarily verifiable.
+
+The remaining two axioms tell us that the finite conjunction and the countable disjunction of verifiable statements is verifiable.
+That is, we can use the tests of the statement to construct the test for the logical AND (i.e.
+check that all tests terminates successfully) and the logical OR (i.e.
+check that at least one test terminates).
+
+These axioms provide the ability to keep track the logical relationships between a set of statements and their connection to experimental evidence.
+
 \section{Experimental and theoretical domains}

### 2021-02-26 Articles/2020-QuantumLogic/Source/quantumlogic.tex
First pass at the math

@@
 \usepackage{amsfonts}
+\usepackage{amsthm}
 \usepackage{MnSymbol}
@@
 \frenchspacing
+
+\newtheorem{prop}[equation]{Proposition}
 
@@
 We also have ``the support of $\rho$ is $U \cup V$'' is not equivalent to ``the support of $\rho$ is $U$'' $\vee$ ``the support of $\rho$ is $V$''.
-The distribution could span both sets at the same time.
+The dis tribution could span both sets at the same time.
 But the lattice generated by such sets is still a $\sigma$-algebra, and therefore follows classical logic.
@@
 \section{Math}
+
+In this section, we want to show how Hilbert spaces already are equipped with a mathematical structure that follows classical logic, namely the Borel algebra.
+We will see that the algebra typically associated with quantum logic is a subset of this structure.
+We will see that this classical structure contains other statements that are experimentally interesting, including statements about the expectation values.
+
+Let us start with a brief summary of the basic mathematical structure of quantum logic.
+Each proposition for a Hilbert space $H$ correspond to a closed subset.
+This forms a complete lattice under the following definitions: the ordering $p \leq q$ of the lattice (i.e.
+the implication $p \to q$) represents subspace inclusion, the conjunction $p \wedge q$ is the biggest subspaces contained by both, the disjunction $p \vee q$ is the smallest subspace that contains both, the negation $\neg p$ is the orthogonal subspace.
+
+To understand how the previous structure fails to a be Boolean algebra, we note that by Stone's representation theorem any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.
+inclusion, intersection, union and complement).
+If we understand each subspace as a set of vector, the conjunction $p \wedge q$ is not the mere set union of the subspaces, but the span.
+This is why the above structure, even if it is a complemented $0-1$ lattice, it fails to be a Boolean alebgra.
+
+However, a Hilbert space $H$ is a complete normed space.
+The norm induces a topology\footnote{The Hilbert space is typically defined to be separable, so the topology will be second countable.}, which in turns induces a Borel-algebra.
+The Borel-algebra is a family of sets closed under countable intersection, union and negation.
+It therefore forms a Boolean algebra.
+Therefore propositions of the type ``the state is in set $A$'' for any Borel set $A$ follow the rules of classical logic.
+
+Note that the lattice of quantum logic is formed by closed subspaces, which are closed sets and therefore Borel sets.
+Therefore all statements expressible in quantum logic (e.g.
+``we measure/prepared this quantity with the following value'') are also expressible with the Borel algebra.
+The converse, however, is not true.
+
+To show that some of these statements are indeed interesting, let us consider the following proposition.
+\begin{prop}
+	Let $H$ be an Hilbert space and $\langle \, , \, \rangle$ be its inner product.
+Let $X$ be a self-adjoint bounded linear operator.
+Let $F_X : H \to \mathbb{R}$ the map defined by $F_X(\psi) = \langle \psi , X \psi \rangle$.
+Let $U \subseteq \mathbb{R}$ be a Borel set.
+Then $F_X^{-1}(U)$ is a Borel set.
+\end{prop}
+
+A proof would work along these lines.
+A linear operator between two normed spaces is continuous if and only if it is bounded, therefore our operator $X$ is continuous.
+Since the inner product is also continuous, $F_X$ is continuous.
+Since $F_X$ is continuous, the reverse image of an open set is an open set.
+Since a Borel set $U$ can be constructed from open sets using negation and countable union, the reverse image $F_X^{-1}(U)$ can also be constructed from open sets using negation and countable union, which means it is a Borel set.
+
+Since $\langle \psi , X \psi \rangle$ expresses the expected value for operator $X$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators.
+However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limit.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity.
+Renormalization would lead to the set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $H \setminus \{ 0 \}$.}
+
+Borel sets allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
+This is NOT equivalent to the statement ``the particle is in the eigenstate $x_0$''.
+The second will be true only if the whole function is at $x_0$.
+That is, if the $x_0$ is the support of the wave-function.
+The first is a lot less stringent: the wave-function can be distributed in infinitely many ways, just as long the average is $x_0$.
+This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions.
+For example, we can take a Gaussian wave packet at center it at any value of position and momentum.
+
+Note that all statements that characterize the position and momentum of the center mass and form a sub-algebra of the Borel algebra.
+Since these are average values, they allow infinitely precise values.
+For example, we can take a Gaussian wave packet at center it at any value of position and momentum.
+Therefore the potential can be considered constant over the wave-function\footnote{In this case, $\langle V(x) \rangle = V(\langle x \rangle)$ and Ehrenfest theorem will reduce to the classical equation.}, the motion of the center of mass will reduce to the case of a point-particle.
+In fact, this is exactly how the point particle approximation arises in classical mechanics as well.
+When we study the motion of planets, cannonballs or specs of dust, we understand that these objects are not point-like: they have an extent and a matter distribution.
+The point-particle approximation works insofar we are interested in the motion of the center of mass and the forces are such that the extent of the object does not matter.
+If the orbit of a planet would be such that its distance is less than the radius of the sun, the point-particle approximation would fail spectacularly.
+
+Now that we have shown that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us do the converse: show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
+As we said before, classical objects are not point-like.
+So let us consider the space of all possible matter distributions $\rho(x,p)$ over phase-space.
+We can formalize this space a Hilbert space as well, where vector $\psi$ corresponds to $\sqrt(\rho(x,p))$.
+This will be a real vector space, and we are only going to consider operators of the form $f(x,p)$.
+Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.
+We can now take the algebra of subspaces and, as in the quantum case, it will be non-distributive: the disjunction of two subspaces will not be the set union, but the span.
+Whether this construction is useful in classical physics is debatable, we find it useful to understand the mathematics and see how it is not a property of quantum mechanics per se: it's a property of vector space.
+
+To simplify, let us assume that there are only two possible states $1$ and $2$.
+A state is identified a distribution $(\rho_1, \rho_2)$, which we can take it to be either the probability of the particle being in each state or the amount of material in each state.
+We note $(k,0)$ the subspace spanned by the first state only.
+This corresponds to the case where the whole system is in state $1$.
+Conversely, $(0,k)$ corresponds to the case where the whole system is in state $2$.
+The disjunction $(k,0) \vee (0,k)$ will correspond to the state spanned by both, which is the whole space $(k,j)$.
+We can also take the subspace $(k,k)$ which is the subspace spanned by vectors that have equal components in $1$ and $2$: the system is equally distributed in the two cases.
+Note that $(k,0)$, $(0,k)$ and $(k,k)$ are all pairwise disjoint: no pair have a vector in common apart from zero.
+Therefore their conjunctions $(k,0) \wedge (0,k) = (0,k) \wedge (k,k) = (k,k) \wedge (k,0) = (0,0)$ are the zero dimensional.
+On the other hand, $(k,k)$ is a subspace of the whole space $(k,j)$ and therefore $(k,k) \wedge (k,j) = (k,k)$.
+Putting it all together, we have $(k,k) \wedge ( (k,0) \vee (0,k) ) = (k,k) \wedge (k,j) = (k,k)$.
+On the other hand, $( (k,k) \wedge (k,0) ) \vee ( (k,k) \wedge (0,k) ) = (0,0) \vee (0,0) = (0,0)$.
+Therefore the lattice is not distributive.
+
+The lack of distributivity has nothing to do with the complexity of the Hilbert space or non-commuting observables.
+It's a property of the lattice of subspaces of a vector spaces.
+In fact, it's a property of the lattice on subgroups more in general.
+One may argue that it is appropriate to focus on this lattice in quantum mechanics while it is not appropriate for classical mechanics.
+But this difference is not a fundamental difference in the mathematical or logical structure of theories.
+
+The difference between the algebras of statements, then, is not really about the type of the logical structure: it is its content in terms of statements.
+If we take two intervals $U_x$ and $U_p$ in position and momentum, and form the statement ``the state is wholly in $U_x$ and $U_p$'', meaning that the vector components are zero in those regions, in the classical case we always have at least $\rho(x,p)$ that satisfies the bounds, while in classical mechanics we do not.
+The logic can be made distributive or non-distributive in either case.
+It is the content of the logic that changes.
+
+\section{Old notes}
 

### 2021-03-01 Articles/2020-QuantumLogic/Source/quantumlogic.tex
Adding math section for quantum logic

--- 2021-02-18 Articles/2020-QuantumLogic/Source/quantumlogic.tex
@@
 \section{Math}
+
+In this section, we want to show how Hilbert spaces already are equipped with a mathematical structure that follows classical logic, namely the Borel algebra.
+We will see that the algebra typically associated with quantum logic is a subset of this structure.
+We will see that this classical structure contains other statements that are experimentally significant, like statements about the expectation values.
+Finally, we will show that the algebra of subspaces of classical distributions is also non-distributive in the same way that quantum logic is.
+
+Let us start with a brief summary of the basic mathematical structure of quantum logic.
+Each proposition for a Hilbert space $H$ correspond to a closed subset.
+This forms a complete lattice under the following definitions: the ordering $p \leq q$ of the lattice (i.e.
+the implication $p \to q$) represents subspace inclusion, the conjunction $p \wedge q$ is the biggest subspaces contained by both, the disjunction $p \vee q$ is the smallest subspace that contains both, the negation $\neg p$ is the orthogonal subspace.
+
+To understand how the previous structure fails to a be Boolean algebra, we note that by Stone's representation theorem any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.
+inclusion, intersection, union and complement).
+If we understand each subspace as a set of vectors, the conjunction $p \vee q$ is not the mere set union of the subspaces, but the span.
+This is why the above structure, even if it is a complemented $0-1$ lattice, it fails to be a Boolean alebgra.
+
+However, a Hilbert space $H$ is a complete normed space.
+The norm induces a topology, which in turns induces a Borel-algebra.
+The Borel-algebra is a family of sets closed under countable intersection, union and negation.
+It therefore forms a Boolean algebra.
+Therefore propositions of the type ``the state is in set $A$'' for any Borel set $A$ follow the rules of classical logic.
+
+Note that the lattice of quantum logic is formed by closed subspaces, which are closed sets and therefore Borel sets.
+Therefore all statements expressible in quantum logic (e.g.
+``we measured/prepared this quantity with the following value'') are also expressible with the Borel algebra.
+The converse, however, is not true: not all Borel sets correspond to propositions in quantum logic.
+
+To show that some of these statements are indeed interesting, let us consider the following result:
+\begin{prop}
+	Let $H$ be an Hilbert space and $\langle \, , \, \rangle$ be its inner product.
+Let $X$ be a self-adjoint bounded linear operator.
+Let $F_X : H \to \mathbb{R}$ the map defined by $F_X(\psi) = \langle \psi , X \psi \rangle$.
+Let $U \subseteq \mathbb{R}$ be a Borel set.
+Then $F_X^{-1}(U)$ is a Borel set.
+\end{prop}
+
+A proof would work along these lines.
+A linear operator between two normed spaces is continuous if and only if it is bounded, therefore our operator $X$ is continuous.
+Since the inner product is also continuous, $F_X$ is continuous.
+Since $F_X$ is continuous, the reverse image of an open set is an open set.
+Since a Borel set $U$ can be constructed from open sets using negation and countable union, the reverse image $F_X^{-1}(U)$ can also be constructed from open sets using negation and countable union, which means it is a Borel set.
+
+Since $\langle \psi , X \psi \rangle$ expresses the expected value for operator $X$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators.
+However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limit.
+Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity.
+Renormalization would lead to the set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $H \setminus \{ 0 \}$.}
+
+Borel sets allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
+This is NOT equivalent to the statement ``the particle is in the eigenstate $x_0$''.
+The second will be true only if the whole function is at $x_0$.
+That is, if the $x_0$ is the support of the wave-function.
+The first is a lot less stringent: the wave-function can even be distributed over an infinite range (e.g.
+a Gaussian wave packet) just as long the average is $x_0$.
+This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.
+a Gaussian wave packet can be centered any value of position and momentum).
+
+Note that all statements that characterize the position and momentum of the center mass and form a sub-algebra of the Borel algebra.
+If the potential for a quantum particle can be considered constant over the wave-function (i.e.
+$\langle V(x) \rangle = V(\langle x \rangle)$ ), the motion of the center of mass will reduce to the case of a point-particle.
+In fact, this is exactly how the point particle approximation arises in classical mechanics as well: when we study the motion of planets, cannonballs or specs of dust, we understand that these objects are not point-like, but they have an extent and a matter distribution.
+The point-particle approximation works insofar we are interested in the motion of the center of mass and the forces are such that the extent of the object does not matter.
+Naturally, if the orbit of a planet would be such that its distance is less than the radius of the sun, the point-particle approximation would fail spectacularly.
+In this regard, quantum mechanics works the same way.
+
+Now that we have shown that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us do the converse: show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
+As we said before, classical objects are not point-like.
+So let us consider the space of all possible matter distributions $\rho(x,p)$ over phase-space.
+We can formalize this as a Hilbert space, where vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$.
+This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$.
+Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.\footnote{To be clear: we are not arguing that this construction is the most appropriate or useful.
+Simply that it can be done.}  We can now take the algebra of subspaces and, as in the quantum case, it will be non-distributive: the disjunction of two subspaces will not be the set union, but the span.
+
+To simplify, let us assume that there are only two possible states $1$ and $2$.
+A state is identified by a two-valued distribution $[\rho_1, \rho_2]$, which we can take it to be either the probability of the particle being in each state or the amount of material in each state.
+We note $[k,0]$ the subspace spanned by the first state only.
+This corresponds to the case where the whole system is in state $1$.
+Conversely, $[0,k]$ corresponds to the case where the whole system is in state $2$.
+The disjunction $[k,0] \vee [0,k]$ will correspond to the state spanned by both, which is the whole space $[k,j]$.
+We can also take the subspace $[k,k]$ which is the subspace spanned by vectors that have equal components in $1$ and $2$: the system is equally distributed in the two cases.
+Note that $[k,0]$, $[0,k]$ and $[k,k]$ are all pairwise disjoint: no pair have a vector in common apart from zero.
+Therefore their conjunctions $[k,0] \wedge [0,k] = [0,k] \wedge [k,k] = [k,k] \wedge [k,0] = [0,0]$ are the zero dimensional.
+On the other hand, $[k,k]$ is a subspace of the whole space $[k,j]$ and therefore $[k,k] \wedge [k,j] = [k,k]$.
+Putting it all together, we have $[k,k] \wedge ( [k,0] \vee [0,k] ) = [k,k] \wedge [k,j] = [k,k]$.
+On the other hand, $( [k,k] \wedge [k,0] ) \vee ( [k,k] \wedge [0,k] ) = [0,0] \vee [0,0] = [0,0]$.
+Therefore the lattice is not distributive.
+
+The lack of distributivity has nothing to do with the complexity of the Hilbert space or non-commuting observables.
+It's a property of the lattice of subspaces of a vector spaces.
+In fact, it's a property of the lattice on subgroups more in general.
+In the subgroup lattice, the disjunction is the span, so doing the disjuction first makes our set bigger during the calculation.
+Naturally, one may still argue that it is appropriate to focus on this lattice in quantum mechanics while it is not appropriate for classical mechanics.
+But one cannot argue that this difference is a fundamental difference in the mathematical or logical structure of theories.
+Certainly, one cannot say that quantum mechanics requires a departure from classical mechanics when both theories follow similar structures.
+
+The key difference between the algebras of statements, then, is not really about the type of the logical structure: it is its content in terms of statements.
+If we take two intervals $U_x$ and $U_p$ in position and momentum, and form the statement ``the state is wholly in $U_x$ and $U_p$'', meaning that the vector components are zero outside of those regions, in the classical case we always find at least one $\rho(x,p)$ that satisfies the bounds, while in quantum mechanics, if the bounds violate the uncertainty principle, no suitable $\psi(x)$ can be found.
+It is the content of the logic that changes.
+
+\section{Old notes}
 

### 2021-03-01 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex

@@
 The intellectual richness of classical mechanics manifests itself in an abundance of formulations, with Newtonian, Hamiltonian, Lagrangian, and Hamilton--Jacobi being among the most prominent.
-Butterfield \parencites*[]{Butterfield2006} has aptly referred to these formulations as different \textit{frameworks} for treating classical systems.
-Recent work in philosophy of science has focused largely on how these formulations relate to one another, with the aim of arriving at a deeper understanding of classical mechanics.
+Butterfield \parencites*[]{Butterfield2004} has aptly referred to these formulations as different \textit{schemes} for treating classical systems.
+Recent work in philosophy of science has focused largely on how these schemes relate to one another, with the aim of arriving at a deeper understanding of classical mechanics.
 By considering the conceptual differences between these formulations, we can better interpret the physical significance of our representations.
 
-In the context of comparing Hamiltonian with Lagrangian mechanics, three interpretive stances have emerged.
+In the context of comparing Hamiltonian and Lagrangian mechanics, three interpretive stances have emerged.
 \textcites[]{North} has argued for privileging Hamiltonian mechanics, on the grounds that it requires less mathematical structure to formulate.
@@
 However, he also shows that if we consider a more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are \textit{not} categorically equivalent \parencites*[1182-83]{Barrett2}.
-Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation, thereby avoiding trivialization worries from categorical equivalence.} Whereas North's argument relies on a problematic criterion for comparing structure---based on the size of symmetry groups \parencites[]{Swanson}---our argument avoids this pitfall.
+Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation.} Whereas North's argument relies on a problematic criterion for comparing structure (based on the size of symmetry groups) \parencites[]{Swanson}, our argument avoids this pitfall.
 We show that a system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian.
@@
 In general, knowing a system's position and velocity does not suffice for knowing its momentum.
-Considering a photon as a classical particle, we cannot determine its momentum from its trajectory through space.\footnote{Since photons travel at constant speed, we can only recover the direction of momentum from the trajectory, but not the magnitude.
-Two photons with different momenta can follow the same trajectory.} Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
+Considering a photon as a classical particle, we cannot determine its momentum from its trajectory through space.\footnote{Since photons travel at constant speed, we can recover the direction of momentum---but not the magnitude---from the trajectory.
+Two photons with different momenta can follow the same trajectory in physical space.} Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
 Different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
@@
 Our derivation of Lagrangian mechanics in Section~\ref{motion} takes place within the context of conservative systems, so we do not consider systems that satisfy inhomogenous Euler--Lagrange equations.
-Although both the Lagrangian and Hamiltonian formulations apply to dissipative systems \parencites[\S 10.4]{Cline}, we exclude these.\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
+Although both the Lagrangian and Hamiltonian formulations apply to dissipative systems, we exclude these.\footnote{For details, see \parencites[\S 10.4]{Cline}.
+Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
 For details of the Lagrangian approach to dissipative systems, see \textcites[]{Smith}.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime.
@@
 
-Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus structure.
+Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus.
 Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration.
@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{Similarly, Butterfield \parencites[]{Butterfieldpoints} discusses point-particles as being extensionless and Wilson \parencites[]{Wilson} discusses them as being zero-dimensional.}
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.
+For further discussion, see \textcites{shpmp}.
+[omit this for review].}
 %comment on how this characterization of infinitesimal particles relates to standard philosophy of physics interpretive disputes over the status of point particles in classical mechanics.
@@
 
-If $\mathcal{S}$ is a discrete space with $U \subseteq \mathcal{S}$, we define $\mu(U) = \sum\limits_{s \in U} \rho_c(s)$ so that $\mu(\mathcal{S})$ gives us the size of the system and $\mu(U)/\mu(\mathcal{S})$ gives us the fraction of the particles that can be found in the region $U$.\footnote{The discrete case is also far less interesting because discrete quantities cannot change continuously over time, meaning that they must be constant.
+If $\mathcal{S}$ is a discrete space with $U \subseteq \mathcal{S}$, we define $\mu(U) = \sum\limits_{s \in U} \rho_c(s)$ so that $\mu(\mathcal{S})$ gives us the size of the system and $\mu(U)/\mu(\mathcal{S})$ gives us the fraction of the particles that can be found in the region $U$.\footnote{The discrete case is far less interesting because discrete quantities cannot change continuously over time, meaning that they must be constant.
 For details see \textcites[]{AoPPhy1}.
@@
 Here, $\xi(s) \equiv \xi^a$ and $\hat{\xi}(s) \equiv \hat{\xi}^b$.} Likewise, the integral $\int_U \rho_c(s) d\mathcal{S} $ is independent of the choice of state variables, so $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.
-By the change of variables formula, we also know that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a= \int_{\hat{\xi}(U)} \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.
-Here, $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ is the determinant of the Jacobian for this change of state variables.
-Combining these two equations, we see that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b = \int_{\hat{\xi}(U)} \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_{c, \hat{\xi}} (\hat{\xi}^b)d \hat{\xi}^b$, entailing that the Jacobian determinant must equal one.
+By the change of variables formula, we also know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
+For convenience, we denote $\rho_{c, \hat{\xi}} ((\hat{\xi} \circ \xi^{-1})(\xi^a))$ by $\rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
+Likewise, in the numerator of the Jacobian, we denote $\partial(\hat{\xi} \circ \xi^{-1})$ by $\partial \hat{\xi}^b$, where here $\hat{\xi}^b$ is understood as a function of $\xi^a$.} Here, $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ is the determinant of the Jacobian for this change of state variables.
+From the preceding two equations, we see that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.
+Matching integrands, we have $\rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$.
+From the scalar transformation property, $\rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
+Hence, we see that $\rho_{c, \hat{\xi}} (\hat{\xi}^b) =  \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$, entailing that the Jacobian determinant must equal one.
 By dimensional homogeneity, it must also be dimensionless.
@@
 In general, the unit variables $q^i$ are the subset of the state variables that define the unit system, which is needed to relate the properties of the system to the measured values.
-The idea is that the $q^i$ not only determine the units for themselves, but for all other state variables.
+The $q^i$ determine the units not only for themselves, but for all other state variables.
 A \textit{change of units} is therefore a map $\hat{q}^j = \hat{q}^j(q^i)$ that defines new unit variables in terms of the old unit variables alone.
@@
 Then, the state variables are given by a set $\xi^a = \{ q, k^\alpha \}$, where \textit{prima facie} the index $\alpha$ could be any non-negative integer.
-We will show that, in fact, there must be exactly one $k^i$ variable, conjugate to the single $q$ variable.
+We will show that, in fact, there must be exactly one $k^\alpha$ variable, conjugate to the single $q$ variable.
 % We further assume that $q$ fully defines the units for all state variables.
@@
 As shown above, the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one.
-Assume for contradiction that there is only one state variable, so that there are no $k^i$ variables.
+Assume for contradiction that there is only one state variable, so that there are no $k^\alpha$ variables.
 Then we would have $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right| = 1$.
@@
 The determinant of a triangular block matrix equals the product of determinants of the block matrices along the diagonal: $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| =  \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^\beta}{\partial k^\alpha}\right|$.
-For the Jacobian determinant to equal one, we must therefore have  $\left|\frac{\partial \hat{k}^b}{\partial k^a} \right| =   \left|\frac{\partial \hat{q}}{\partial q}\right|^{-1}$.
+For the Jacobian determinant to equal one, we must therefore have  $\left|\frac{\partial \hat{k}^\beta}{\partial k^\alpha} \right| =   \left|\frac{\partial \hat{q}}{\partial q}\right|^{-1}$.
 Since $\hat q$ is a function only of $q$, $\frac{\partial \hat{q}}{\partial q}$ is a total derivative, so we obtain the inverse by swapping the numerator and denominator.
@@
 Consequently, there must be exactly two state variables: $q$ and a single $k$.
-So in the case of a single $q$ variable, defining a coordinate-independent density requires a two-dimensional manifold.
+So in the case of a single $q$ variable, defining an invariant density $\rho_{c, \xi}$ requires a two-dimensional manifold.
 
@@
 This agrees with the practice in statistical mechanics.
-The density $\rho_{c, \xi}$, when normalized, is given in terms of the inverse of the units for $\omega$.
+When normalized, the density $\rho_{c, \xi}$ is given in terms of the inverse of the units for $\omega$.
 In accordance with the standard conventions in quantum mechanics and statistical mechanics, the $k_i$ are given in terms of inverse units of the respective $q^i$, such that $dq^i \wedge dk_i$ is a pure number.
@@
 
-In terms of physics, then, a set of canonical coordinates is a set of state variables that express the density in the correct units.
+We can thereby understand canonical coordinates as a set of state variables that express the density in the correct units.
 A canonical transformation is one that preserves the unit and the numerical value of the density.
@@
 
-In our case, this applies to both the state of the whole system and to the state of its parts.
+In our case, this applies to both the state of the whole system and the state of its parts.
 Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
-Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$, at the initial time $t_0$, uniquely identifies the state of the system.
+Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$---at the initial time $t_0$---uniquely identifies the state of the system.
 Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial state at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e.
-the density remains the same throughout the evolution.
-In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa.
+the density remains the same throughout the evolution.\footnote{For convenience, we will sometimes drop the composite state $c$ and coordinate-choice $\xi$ indices from the invariant density $\rho_{c, \xi} (\xi^a)$.
+The fact that the density is not coordinate-free is central to our argument in Section~\ref{density}.
+We will reserve the word `distribution' for the coordinate-free distribution $\rho_c (s)$.
+} In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa.
 %[could two different particles occupy the same initial state?].
@@
 trajectories in physical space-time) suffices to recover its dynamics (i.e.
-evolutions in state space) and vice-versa.\footnote{ \textcites[218]{Abraham}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time (see \textcites[1180-1181]{Barrett2} for discussion).
+evolutions in state space) and vice-versa.\footnote{ \textcites[218]{Abraham}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time; see \textcites[1180-1181]{Barrett2} for discussion.
 Motion equivalence can be understood as the converse of this theorem: assuming that Lagrangian and Hamiltonian mechanics agree on particle evolutions, we derive that the Legendre transformation is well-defined in the hyper-regular domain.}
@@
 If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.
-Note that each space variable $x^i$ is a unit variable, i.e.
+Note that each space variable $x^i$ is a unit-variable, i.e.
 a state variable that defines a unit.
-In fact, the trajectories can be fully described only by these unit variables.
+In fact, the trajectories can be fully described only by these unit-variables.
 For, to specify a trajectory, we need epistemic access to a unit system, where the number of independent units is determined by the dimensionality of physical space.
@@
 Since these trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
-If the equations of motion were such that the velocity were a function of only the unit variables (i.e.
+If the equations of motion were such that the velocity were a function of only the unit-variables (i.e.
 $v^i=v^i(q^i)$), then motion equivalence would fail.
@@
 Let $H$ denote the Hamiltonian.
-Since by Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
-Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere, which we take to be positive.
+By Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
+Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere---which we take to be positive.
 Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables.
@@
 distributions on a subset of the unit variables), entailing that all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
-We set:\footnote{In Section~\ref{dependence}, we will see that $g^{ij}$ corresponds to the Riemannian metric, since $\partial v^i/\partial p_j$ it is a function only of position.}
+We set:\footnote{In Section~\ref{dependence}, we will see that $g^{ij}$ corresponds to the Riemannian metric, since $\partial v^i/\partial p_j$ is a function only of position.}
 \begin{equation}
@@
 where V is another arbitrary function.
-We recognize this as the Hamiltonian for massive particles under potential forces, covering both classical electrodynamics and classical Newtonian gravitation.
+We recognize this as the Hamiltonian for massive particles under potential forces, relevant for classical electrodynamics and Newtonian gravitation.
 For example, $A$ can play the role of the magnetic vector potential, with $V$ playing the role of the electric potential.
@@
 Section~\ref{fundamental} argues that the invariant phase space density is more fundamental than the state variables, since conjugate momentum is a gauge-dependent quantity.
-We then turn in Section~\ref{Lagrangian} to provide a geometric interpretation of the principle of stationary action within Hamiltonian mechanics.
-Our construction makes manifest the gauge redundancy of the Lagrangian, providing a further reason to privilege the Hamiltonian.
+In Section~\ref{Lagrangian} we provide a geometric interpretation of the principle of stationary action within Hamiltonian mechanics.
+Our construction makes manifest the gauge-dependency of the Lagrangian, providing a further reason to privilege the Hamiltonian.
 Finally, we respond to Curiel's arguments against the fundamentality of Hamiltonian mechanics.
@@
 Additionally, we argue that viewing physics through the lens of invariant densities has much to commend it.
-In general, it provides a better understanding of classical systems.
+In general, it provides a better understanding of classical systems.\footnote{It also provides a better understanding of statistical mechanics and entropy, since phase space provides a uniform measure over q and p.
+For discussion, see \textcites{shpmp} [omit for review].}
 
@@
 This relationship shows that the units of $\rho$ depend on the units of both $m$ and $V$.
+If $\rho(x, y)$ is a function of multiple variables, dimensional analysis requires that the units of $\rho$ are consistent with the units of the arguments in the expression $\int_U \rho(x, y) dx dy$.
 The issue is that a manifold, in general, does not provide us with a way to measure areas and volumes.
-If $\rho(x, y)$ is a function of multiple variables, dimensional analysis requires that the units of $\rho$ are consistent with the units of the arguments in the expression $\int_U \rho(x, y) dx dy$.
 For example, a manifold may capture the physical characteristics of a group of people in terms of their height $h$ and their weight $w$, leading to a density $\rho(h,w)$ with units of people/(m$\cdot$Kg).
@@
 Hence, the density is coordinate-dependent because its definition requires us to make a choice of unit-variables.
-Worse than that, it is not generically coordinate invariant, since the density changes with the Jacobian.
+Worse than that, it is not generically coordinate-invariant, since the density changes with the Jacobian.
 In the above example, $\rho(h,w) = |J| \rho(h, bmi) = \frac{1}{h^2} \rho(h, bmi)$.
@@
 This makes the value of the density, in general, not directly comparable at different points.
-Without additional structure, there is no objective sense in which the density is higher at one point or another: it depends on the coordinate.
+Without additional structure, there is no objective sense in which the density is higher at one point or another: it depends on the coordinates.
 
@@
 
-The Hamiltonian formalism fixes this issue by assigning areas to regions in a way that is independent of the choice of unit variables.
+The Hamiltonian formalism fixes this issue by assigning areas to regions in a way that is independent of the choice of unit-variables.
 To do so, it forces us to restrict ourselves to canonical coordinates, which explains why momentum has to change covariantly.
@@
 In other words, the metric tensor defines a unit for the volume form (e.g.
-meters cubed), and we always express the volume in those units, regardless of what coordinates we picked.
-That would be the natural choice of volume in a Cartesian system, for example, but not in a spherical one.
+meters-cubed), and we always express the volume in those units, regardless of what coordinates we chose.
+This would be a natural choice of volume in a Cartesian system, for example, but not in a spherical one.
 %That is: to This is only possible if the determinant of $g_{ab}$ equals one. Hence, we must define $\hat{\rho} $ in a special coordinate system, namely one where $|g_{ab}| = 1$. 
@@
 In polar coordinates, the natural choice of units would be meter $\cdot$ radians.
-However, the volume form will still return meters squared, since we will have $|g_{ab}| = r^2$ and $dV = r dr \wedge d\theta$.
+However, the volume form will still return meters-squared, since we will have $|g_{ab}| = r^2$ and $dV = r dr \wedge d\theta$.
 This means that $\hat{\rho}$ at best expresses the density in a particularly convenient Cartesian-like coordinate system: one in which the volume is expressed in special units and we can compare the density point-wise.\footnote{This is similar to what happens with the concept of proper time: it is defined in the co-moving frame, so it is coordinate-dependent.
@@
 The first is epistemic: by representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
-Any and all (time-independent) coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
+Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
 See Section~\ref{gauge} for discussion.} Not only is this formal property methodologically advantageous, but it also bespeaks a further physical significance.
@@
 
-Section~\ref{infinitesimal} demonstrated a third reason for viewing the density as fundamental: by defining the density, we recover the right mathematical structure, namely a symplectic manifold.
+Section~\ref{infinitesimal} demonstrated a third reason for viewing the density as fundamental: by defining the density $\rho_{c, \xi}$, we recover the right mathematical structure, namely a symplectic manifold.
 If we did not define the density, then we would need another account of why this mathematical structure arises for classical systems.
@@
 
-Proponents of this necessary condition might respond by noting that although the density is not coordinate-independent, the distribution is.
+Proponents of this necessary condition might respond by noting that although the density $\rho_{c, \xi}$ is not coordinate-independent, the distribution $\rho_{c}$ is.
 For the distribution is equivalent to the classical state, which is a thing in the world, i.e.
@@
 However, we maintain that the density itself is of fundamental significance, both epistemically and physically.
-For it is only through the density that we gain epistemic access to the distribution.
-Although the choice of coordinates is fairly arbitrary, we need to make some choice in order to quantitatively measure the distribution through experiments.
+For it is only through the density $\rho_{c, \xi}$ that we gain epistemic access to the distribution.
+Although the choice of coordinates $\xi$ is fairly arbitrary, we need to make some choice in order to quantitatively measure the distribution $\rho_{c}$ through experiments.
 Additionally, defining the density is necessary for deriving symplectic structure, as shown in Section~\ref{infinitesimal}.
@@
 
-[XXX To do: insert discussion of statistical mechanics and entropy here: statistical mechanics works best on phase space because phase space gives us a uniform measure over q and p.
+%[XXX To do: insert discussion of statistical mechanics and entropy here: statistical mechanics works best on phase space because phase space gives us a uniform measure over q and p.
 Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more salient/perspicuous.
@@
 This is related to the fact that, in Lagrangian mechanics, Cartesian coordinates are a physically significant coordinate choice because they decompose position and velocity variables as independent degrees of freedom.
-This independence does not hold generically across coordinate systems.\footnote{In contrast, canonical position and conjugate momentum variables always characterize separable/independent degrees of freedom in the Hamiltonian framework.} Finally, as the derivation in Section~\ref{deterministic} shows, a necessary condition for defining the symplectic structure is that we can express the density in a particular choice of units.
+This independence does not hold generically across coordinate systems.\footnote{In contrast, canonical position and conjugate momentum variables always characterize separable/independent degrees of freedom in the Hamiltonian framework.} In a similar vein, the next section shows that the conjugate momentum is gauge-dependent, but it is presumably still physically significant.
+%Realized that the following remarks probably go with earlier argument about coordinate-dependence: Finally, as the derivation in Section~\ref{infinitesimal} shows, a necessary condition for defining the symplectic structure is that we can express the density $\rho_{c, \xi}$ in a particular choice of units.
 If we did not have the physical ability to define the density in some units we would not be able to define this structure.
 This provides another instance of a physically significant but coordinate-dependent representational choice.
-Similarly, the next section shows that the conjugate momentum is gauge-dependent, but presumably still physically significant.
 
@@
 
-Before proceeding, we want to give a precise account of what gauge redundancy physically represents.
-In the spirit of our previous constructions, we want to find a physically-motivated definition that leads to the correct mathematical structure.
-We first note that, when motion equivalence holds, we can completely characterize states in terms of the position $q$ and its first derivative $\dot{q}$.
+Before proceeding, we provide a precise account of what gauge-dependency physically represents.
+In the spirit of our previous constructions, we consider a physically-motivated definition that leads to the correct mathematical structure.
+First, note that when motion equivalence holds, we can completely characterize states in terms of the position $q$ and its first derivative $\dot{q}$.
 Hence, any canonical transformation in phase space that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant.
@@
 
-Additionally, the gauge-dependency of conjugate momentum shows that not all state variables are on equal physical footing.
+Additionally, the gauge-dependency of conjugate momentum shows that all state variables are \textit{not} on equal physical footing.
 Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.
@@
 This raises a number of objections to our own position, which we rebut.
-In short, we  argue that we have given an eminently natural construction of Hamiltonian mechanics.
+In short, we  argue that we have given a natural construction of Hamiltonian mechanics.
 %In short, we argue that we have given an eminently natural construction of Lagrangian mechanics, viewed as a special case of Hamiltonian mechanics.
@@
 Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition.
-We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a similar argument, see \textcites[236--37, 243--45]{Arnold}.}
+We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a related argument, see \textcites[236--37, 243--45]{Arnold}.}
 %our derivation seems very similar to Arnold chapter 9, section 45:C: ``the principle of least action in phase space''. Note that we briefly discussed Arnold on 11-23-20 meeting, around 12:30. 
@@
 
-Treating time $t$ as both a coordinate and the parameter of the trajectory, the time component of the displacement, $S_t = dt/ dt = 1$.\footnote{Considering $dt/dt$, the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
+Treating time $t$ as both a coordinate and the evolution parameter, the time component of the displacement $S_t = dt/ dt = 1$.\footnote{In the expression `$dt/dt$', the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
 We could instead use a parameter $\tau$, and the derivation would remain valid.} This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $.
@@
 Integrating this expression, we find that $\theta_q = p$ up to addition by an arbitrary constant function $g(q)$.
-Again using the gauge freedom of $\vector{S}$ we can set $g(q)=0$, so that $\theta_q = p$.
+Again using the gauge freedom of $\vector{S}$, we can set $g(q)=0$, so that $\theta_q = p$.
 Thus, so far, we have shown that $\vector{\theta} = (p, 0,\theta_t) $.
@@
 Note that $\vector{S}  dt = (\frac{d q }{d t } dt, \frac{d p }{d t } dt, \frac{d t }{d t }dt ) = (dq, dp, dt) = d\vector{\gamma}$ represents the infinitesimal displacement along the trajectory.
-Therefore the action for those trajectories can be rewritten as the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.
+Therefore, the action for those trajectories can be rewritten as the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.
 To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
@@
 Hence, by Stoke's theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot dC = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ .
-Recalling now that 1) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and 2)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $\vector{\theta}$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
-This establishes Hamilton's principle of stationary action, which we have derived by assuming that 1) the displacement vector is divergenceless and 2) using the Legendre transform (assuming motion equivalence).
+Recalling now that i) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and ii)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $curl(\vector{\theta})$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
+This establishes Hamilton's principle of stationary action, which we have derived by assuming that 1) the displacement vector is divergenceless and 2) using the Legendre transform (well-defined when motion equivalence holds).
 
@@
 Despite this advantage, Lagrangian mechanics has a compensating disadvantage: the Lagrangian is itself gauge-dependent.
-The formalism of Section~\ref{action} makes this gauge dependence manifest.
+The formalism of Section~\ref{action} makes this gauge-dependence manifest.
 Representing the Lagrangian as $L =\vector{\theta} \cdot \vector{S} $, we see that since the vector potential $\vector{\theta}$ is gauge-dependent while the displacement vector $\vector{S}$ is not, $L$ must also be gauge-dependent.\footnote{As shown in Section~\ref{action} we can arbitrarily transform $\vector{\theta}$ by adding the gradient of an arbitrary scalar function, while $\vector{S}=curl(-\vector{\theta}) $ remains invariant under this gauge transformation.} The gauge-dependence of the Lagrangian makes it difficult to provide a clear physical interpretation of any particular Lagrangian.\footnote{The action inherits the gauge-dependence of the Lagrangian.
@@
 
-Nonetheless, reformulating Hamiltonian mechanics within the extended phase space $(q_i, p_i, t, E)$ provides one way to rectify these deficiencies.\footnote{For a detailed discussion of the extended phase space formalism, see \textcites[]{Struckmeier}.} At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and a time-like parameter within the usual Hamiltonian framework.
-Yet, differential geometry demands that we separate these two roles.
-We thereby add a temporal degree of freedom $(t, E) $, expanding the phase space degrees of freedom by one (recall that we previously had one degree of freedom for every canonical coordinate pair $q_i, p_i$).
+Nonetheless, reformulating Hamiltonian mechanics within the extended phase space $(q^i, p_i, t, E)$ provides one way to rectify these deficiencies.\footnote{For a detailed discussion of the extended phase space formalism, see \textcites[]{Struckmeier}.} At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and a time-like parameter within the usual Hamiltonian framework.
+Yet, differential geometry recommends that we separate these two roles.
+We thereby add a temporal degree of freedom $(t, E) $, expanding the phase space degrees of freedom by one (recall that we previously had one degree of freedom for every canonical coordinate pair $q^i, p_i$).
 Next, we define the phase space evolution over an evolution parameter $s $.
-This yields two more equations of motion: $dt/ds = -\partial \mathcal{H}/ \partial E$ and $dE/ds = -\partial \mathcal{H}/ \partial t$, where $\mathcal{H}(q_i, p_i, t, E)$ is the extended Hamiltonian.
+This yields two more equations of motion: $dt/ds = -\partial \mathcal{H}/ \partial E$ and $dE/ds = -\partial \mathcal{H}/ \partial t$, where $\mathcal{H}(q^i, p_i, t, E)$ is the extended Hamiltonian.
 The extended Hamiltonian not only acts as a generator of the evolution parameter $s$, but it also provides us with the \textit{Hamiltonian constraint} $\mathcal{H}=0$.
@@
 He also uses the almost-tangent structure to give a general characterization of what he takes to be the fundamental kinematical constraints.} According to Curiel, all physically-meaningful classical systems satisfy a certain set of kinematic constraints.
-In the simplest case, these constraints characterize velocity as the first temporal derivative of position: $v^i = d_t x^i = v^i(q^j, p_k)$ (equation~\ref{weak_equivalence}).\footnote{Curiel treats this constraint as `kinematical' in the sense that it must hold independently of the interactions between a system and its environment.
+In the simplest case, these constraints characterize velocity as the first temporal derivative of position: $v^i = d_t x^i = v^i(q^j, p_k)$ (Equation~\ref{weak_equivalence}).\footnote{Curiel treats this constraint as `kinematical' in the sense that it must hold independently of the interactions between a system and its environment.
 He interprets it as a substantive physical condition, rather than an analytic truth \parencites*[282]{Curiel}.} Curiel argues that Lagrangian mechanics encodes the necessary and sufficient conditions for a system to satisfy these kinematic constraints, both within and outside the hyper-regular domain \parencites*[307--308, 311]{Curiel}.
@@
 However, we believe that there are physically-meaningful non-regular Hamiltonian systems.
-These include classical models of a massless particle moving at a constant speed, such as a classical photon.
+These include classical models of a massless particle moving at constant speed, such as a classical photon.
 In general, Hamiltonian systems outside the hyper-regular domain correspond to cases where we cannot express the density of states in terms of position and velocity.
@@
 There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.} For instance, a zero Hamiltonian describes a static system (by Hamilton's equations, the derivatives of the position and momenta are zero).
-Indeed, Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}.
+Indeed, Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside the hyper-regular domain \parencites*[291, 308]{Curiel}.
 Curiel argues that this is an advantage of Lagrangian mechanics, but it generates challenges for understanding non-hyper-regular Lagrangians.\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.
@@
 Nevertheless, our definition of classical systems has substantive advantages over Curiel's.
-Preceding sections have already illustrated many of these advantages.
+Preceding sections have illustrated many of these advantages.
 Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{density}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a more geometrical understanding of the principle of stationary action (Section~\ref{action}).
 Additionally, our definition accommodates models of classical massless particles by maintaining the uniqueness of solutions, even outside the hyper-regular domain.
-Additionally, our definition of a classical system has the advantage of an elegant relationship with quantum mechanical systems.\footnote{For details of the quantum case, see \textcites{AoPPhy1}.
-[XXX: omit this for review].} By changing the requirement of infinitesimal reducibility (Section~\ref{infinitesimal}) to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.
+Moving beyond classical mechanics, our framework has the advantage of providing an elegant relationship with quantum mechanical systems.\footnote{For details of the quantum case, see \textcites{AoPPhy1}.
+[omit for review].} By changing the requirement of infinitesimal reducibility (Section~\ref{infinitesimal}) to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.
 Hence, our definition perspicuously distinguishes quantum from classical systems.
@@
 Besides this, our approach has many additional interpretive payoffs.
-Our formalism emphasizes the fundamental significance of the invariant phase space density.
+Our formalism emphasizes the fundamental significance of the invariant phase space density, along with the role of the symplectic form in counting states for each degree of freedom.
 Whereas the state variables $(q, p)$ suffer from the gauge-dependency of the conjugate momentum, the density provides an invariant description within the Hamiltonian formalism.
-This explains why statistical mechanics is best done on phase space, since it provides the natural setting for an invariant density.
-We have also elucidated the role of the symplectic form in counting states for each degree of freedom.
 The importance of the density shows that a quantity can be fundamental yet coordinate-dependent.
 We have also argued that non-fundamental quantities can be physically significant despite being neither coordinate- nor gauge-invariant.
-Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain.
-This treatment makes the gauge redundancy of the Lagrangian manifest.
+Finally, we have provided a geometrical interpretation of the principle of stationary action that makes the gauge-dependency of the Lagrangian manifest.
 Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but classical systems in general.
+%Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain.
+%In fact, this explains why statistical mechanics is best done on phase space, since it provides the natural setting for an invariant density.
+[include only if space to explain this in part of section 5.]
+%We have also elucidated the role of the symplectic form in counting states for each degree of freedom.
 

### 2021-03-01 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Minor changes during the meeting

@@
 Here, $\xi(s) \equiv \xi^a$ and $\hat{\xi}(s) \equiv \hat{\xi}^b$.} Likewise, the integral $\int_U \rho_c(s) d\mathcal{S} $ is independent of the choice of state variables, so $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.
-By the change of variables formula, we also know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
-For convenience, we denote $\rho_{c, \hat{\xi}} ((\hat{\xi} \circ \xi^{-1})(\xi^a))$ by $\rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
-Likewise, in the numerator of the Jacobian, we denote $\partial(\hat{\xi} \circ \xi^{-1})$ by $\partial \hat{\xi}^b$, where here $\hat{\xi}^b$ is understood as a function of $\xi^a$.} Here, $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ is the determinant of the Jacobian for this change of state variables.
+By the change of variables formula, we also know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\phi = \hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
+For convenience, we denote $\rho_{c, \hat{\xi}} (\phi(\xi^a))$ by $\rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
+Likewise, in the numerator of the Jacobian, we denote $\partial \phi$ by $\partial \hat{\xi}^b$, where here $\hat{\xi}^b$ is understood as a function of $\xi^a$.} Here, $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ is the determinant of the Jacobian for this change of state variables.
 From the preceding two equations, we see that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.
@@
 \end{equation}
-where $A_j$ are arbitrary functions. Note that:
+where $A_j$ are arbitrary functions. We recognize $m g_{ji} v^i$ as kinetic momentum. Note that:
 \begin{equation}

### 2021-03-02 Articles/2020-QuantumLogic/Source/quantumlogic.tex
Fixed a few typos

@@
 Since if I measured spin $y$ I would either get $+$ or $-$, we have $q \vee r \equiv \top$.
-Therefore $q \vee r \equiv \top$ is true.
+Therefore $p \wedge (q \vee r)$ is true.
 On the other hand, both terms in the second disjunctions are false: because $p \wedge  q$ refer to incompatible observables, we have $p \wedge  q \equiv \bot$.
@@
 
-$q \vee r \equiv \top$ therefore $p \wedge (q \vee r) \equiv p \wedge \top \equiv p$. If $p$ is true
+\subsubsection{Reframing}
 
-\subsubsection{Reframing}
+2 casi: considerare
 
@@
 Therefore $p \wedge (q \vee r) \equiv \bot \equiv (p \wedge  q) \vee (p \wedge r)$.
-
-2 casi: considerare
 
@@
 Let us start with a brief summary of the basic mathematical structure of quantum logic.
-Each proposition for a Hilbert space $H$ correspond to a closed subset.
+Each proposition for a Hilbert space $H$ correspond to a closed subspace.
 This forms a complete lattice under the following definitions: the ordering $p \leq q$ of the lattice (i.e.
@@
 inclusion, intersection, union and complement).
-If we understand each subspace as a set of vectors, the conjunction $p \vee q$ is not the mere set union of the subspaces, but the span.
+If we understand each subspace as a set of vectors, the disjunction $p \vee q$ is not the mere set union of the subspaces, but the span.
 This is why the above structure, even if it is a complemented $0-1$ lattice, it fails to be a Boolean alebgra.
@@
 Since $\langle \psi , X \psi \rangle$ expresses the expected value for operator $X$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators.
-However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limit.
+However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limited.
 Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity.
@@
 This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.
-a Gaussian wave packet can be centered any value of position and momentum).
+a Gaussian wave packet can be centered on any value of position and momentum).
 
-Note that all statements that characterize the position and momentum of the center mass and form a sub-algebra of the Borel algebra.
+Note that all statements that characterize the position and momentum of the center of mass form a sub-algebra of the Borel algebra.
 If the potential for a quantum particle can be considered constant over the wave-function (i.e.
@@
 So let us consider the space of all possible matter distributions $\rho(x,p)$ over phase-space.
-We can formalize this as a Hilbert space, where vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$.
+We can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$.
 This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$.
@@
 Naturally, one may still argue that it is appropriate to focus on this lattice in quantum mechanics while it is not appropriate for classical mechanics.
-But one cannot argue that this difference is a fundamental difference in the mathematical or logical structure of theories.
+But one cannot argue that this difference is a fundamental difference in the mathematical or logical structure of the theories.
 Certainly, one cannot say that quantum mechanics requires a departure from classical mechanics when both theories follow similar structures.

### 2021-03-04 Overviews/Experimental verifiability/ExperimentalVerifiabilityOverview.tex
First draft at experimental verifiability overview

--- 2021-02-22 Overviews/Experimental verifiability/ExperimentalVerifiabilityOverview.tex
@@
 \frenchspacing
+
+\newcommand\partitle[1]{\textsc{#1}.}
 
@@
 
-The set of points of a topological space represent the possible cases (e.g.
+The set of points of a topological space represents the possible cases (e.g.
 all possible values for the mass of a particle).
@@
 
+\partitle{Principle of scientific objectivity} The guiding principle is that \emph{science is universal, non-contradictory and evidence based.} This means that any scientific theory must specify a set of \textbf{statements} that are logically consistent, whose truth can be confirmed experimentally by everybody.
+This starting point is codified in the axioms that form the formal foundation of the theory.
+
+\partitle{Axioms of logic} The first set of axioms ensure the theory is universal and non-contradictory.
+The \emph{axiom of context} tells us that statements are organized into a \textbf{logical context} $\logCtx$, and a well defined truth value for each statement must exist.
+This recognizes that the logical consistency and meaning of each statement relies on definitions and relationships that depends on the context\footnote{For example, if one is performing particle identification, the mass of an electron is assumed to be given.
+If one is measuring the mass, the identify is assumed to be given, which the value of the mass is unknown.}, on a set of statements.
+
+The \emph{axiom of possibility} tell us that the statements in a context cannot be assigned arbitrary truth values, but only in a way that is consistent with the meaning defined by the context.\footnote{For example, \statement{the mass of the object is one kilogram} and \statement{the mass of the object is two kilograms} cannot be both true.} Therefore each context comes with a set $\pAss$ of possible assignments, where each possible assignment is a function $a : \logCtx \to \Bool$ that returns a value of truth for each statement.
+The truth is a possible assignment.
+
+The \emph{axiom of closure} tells us that we can always find a statements whose truth arbitrarily depends on the truth of other statements.
+
+\partitle{Logical relationships} The logic axioms entail that each logical context is a complete Boolean algebra.
+They allow us to define useful relationship such as: equivalence $\equiv$ (i.e.
+statements that must be equally true or false); narrowness (i.e.
+a statement is more specific than another)\footnote{For example, \statement{the mass of the object is between 4 and 5 kilograms} is narrower than \statement{the mass of the object is between 0 and 100 kilograms}.}; compatibility $\comp$ (i.e.
+whether two statement can be true at the same time); independence $\indep$ (i.e.
+whether the truth of some statements constrains the truth of others).
+This basic structure constrains any further mathematical structure as the two will have to interact appropriately.
+
+\partitle{Axioms of verifiability} This set of axioms capture the rules of experimental verifiability.
+The \emph{axiom of verifiability} tells us that some statements are \textbf{verifiable}, meaning we have a test at our disposal that will terminate successfully in finite time if and only if the statement is true.
+Note that the test is not guaranteed to terminate, to give an answer, if the test is false.\footnote{For example, a test for \statement{there exists extra-terrestrial life} may terminate successfully if life is found, but it will not be conclusive if life is not found.
+On the other hand, \statement{the mass of the particle is exactly 0 eV} can be experimentally disproved, but can never be confirmed due to finite precision.} Therefore the negation of a verifiable statement is not necessarily verifiable.
+It also tells us that a statement equivalent to a verifiable statement is verifiable, as we can use the test of the second to test the first.
+
+The remaining two axioms tell us that the finite conjunction and the countable disjunction of verifiable statements is verifiable.
+That is, we can use the tests of the statement to construct the test for the logical AND (i.e.
+check that all tests terminates successfully) and the logical OR (i.e.
+check that at least one test terminates).
+The disjunction can be extended to a countable set because we only need one test to terminate successfully in a finite time, not all of them.
+
+\partitle{Decidability} A \textbf{falsifiable} statement is one whose negation is falsifiable.
+A \textbf{decidable} statement is one that is both verifiable and falsifiable.
+The finite conjunction/disjunction and negation of decidable statements are decidable.
+
+\partitle{Different algebras} The logical context as a whole, the set of verifiable statements and the set of decidable statements follow different algebras according to table \ref{algebras}.
+
+\begin{table}[h]
+	\centering
+	\begin{tabular}{p{0.10\textwidth} p{0.05\textwidth} p{0.10\textwidth} p{0.10\textwidth} p{0.10\textwidth}}
+		Operator & Gate & Statement & Verifiable Statement & Decidable Statement  \\ 
+		\hline 
+		Negation & NOT & allowed & disallowed & allowed \\ 
+		Conjunction & AND & arbitrary  & finite & finite \\ 
+		Disjunction & OR & arbitrary  & countable & finite \\ 
+	\end{tabular}
+	\caption{Comparing algebras of statements.}
+	\label{algebras}
+\end{table}
+
 \section{Experimental and theoretical domains}
 
-\section{Topologies and $\sigma$-algebras}
+\partitle{Experimental domains} The next step is to group together verifiable statements that characterize the same physical object (e.g.
+all the statements that describe the state of a system or the properties of an object).
+We call such a group an \textbf{experimental domain} $\edomain$ and it will have at least two characteristics.
+First of all, it will be closed under finite conjunction (i.e.
+logical AND) and countable disjunction (i.e.
+logical OR) since these are exactly the operations that give us other verifiable statements.
+Second, it cannot be too big: we must be able to explore it fully given an arbitrarily long amount of time.
+Formally, it can be generated from a countable subset of verifiable statements through finite conjunction and countable disjunction, it must have a countable \textbf{base}.
+This means the tests for all statements in the domain can be generated from the tests of the base, therefore running those countably many tests is enough to test the whole domain.
 
-\section{Experimental relationships and continuous functions}
+\partitle{Theoretical domains} While the verifiable statements define all the testing procedures that are available in some domain, they are not the only valid statements in a scientific theory.\footnote{For example, \statement{the mass of the photon is exactly 0 eV} is not something we can verify due to the limited precision of our statements, yet it is something we want to be able to say in our theory.} We are interested in all statement that can be assigned an experimental test, regardless of its terminating conditions.
+We call these \textbf{theoretical statements} and we group them into a \textbf{theoretical domain} $\tdomain$.
+This is generated from experimental domain by closing under negation and countable disjunction.
+
+\partitle{Possibilities} Within a theoretical domain we find the \textbf{possibilities} $X$ of the domain $\edomain_X$.
+These are statements that give the narrowest description possible and, if assumed to be true, tell us the truth value for all other statements.
+The set of possibilities represent all the possible cases that can be distinguished by the experimental domain.\footnote{For example, if the domain characterizes the state of a bead along the wire, ``the position of the bead is 3 meters and the velocity is zero meters per second'' is a possibility because it characterizes the truth value of all possible experimentally verifiable finite precision descriptions.}
+
+\partitle{Correspondence to topologies and $\sigma$-algebras} Every statement in the domain can be characterize by the set of possibilities for which the statement is true.
+Therefore, we have a correspondence between sets and statement that preserves the structure.
+Under this correspondence, the experimental domain induces a topology where each open set represents a verifiable statement, while the theoretical domain induces a $\sigma$-algebra where each Borel set is a theoretical statement.
+The topology is at least $T_0$ and second countable while the Borel algebra is countably generated (and therefore separable), which make them extremely well-behaved.
+
+\partitle{On the significance of mathematical structures} \textbf{The existence of well-behaved topologies and $\sigma$-algebra is a direct consequence of the requirement of experimental verifiabiliy.}\footnote{Philosophically, one may say that it is not an ontological attribute of the system being studied, but rather a characterization of its epistemically accessible characteristics.
+There is no reason to assume that ontological entities would follow such constrain.} Another important aspect is that \textbf{the points are not the starting point, but rather they are generated from the verifiable statements.} If we change what is being characterized (e.g.
+add measurements of spin to those of the position and momentum of an electron) the points of the space will change.\footnote{Philosophically, the points themselves cannot be understood as some complete ontological entity, but rather the epistemically accessible characteristics under a set of given circumstances.}
+
+\partitle{Unphysical statements} Note that the theoretical domain is closed under countable operations, while the context is closed under arbitrary operations.
+This means that there may be statements that depend on the experimental ones that are not theoretical, they are not associated to a testing procedure: they are unphysical.
+In fact, if the possibilities are uncountable (e.g.
+we use real numbers to identify them, as we do when using position for states) such statements are guaranteed to exist.
+These correspond to non-Borel sets, which therefore have no physical meaning.
+
+\partitle{Maximum cardinality for the possibilities.} Since the domain is generated by a countable set of verifiable statements, the possibilities must be distinguished with a countable sequence of boolean values.
+This means that the cardinality of the possibilities cannot exceed that of the continuum (an infinite sequence of zeros and ones is a real number expressed in base 2), therefore infinities of higher order and all associated problems do not play a role in science.
+
+\partitle{Verifiable, falsifiable and undecidable parts} Given each theoretical statement, we can define its verifiable part (i.e.
+the set of possibilities for which its test terminates successfully), its falsifiable part (i.e.
+test terminates successfully) and the undecidable part (i.e.
+the test does not terminate).
+Topologically, these correspond to the interior, exterior and boundary of the corresponding Borel set.
+\footnote{Note there are domains that contain undecidable statements, those for which the test never terminates.
+For example, \statement{the mass of the proton is rational in units of eV} is undecidable: any finite precision measurement will be consistent with infinitely many rationals and irrational possibilities.
+Mathematically, the boundary of the rational is all the reals.}
+
+\partitle{Discreteness is decidability} A decidable statement will correspond to a clopen set.
+A topology is discrete if all sets are open, which means all the statements are decidable.\footnote{Many physicists confuse discreteness with countability, and characterize the rationals as discrete.
+This depends on how the rationals are measured.
+If the numerator and denominator are measured independently, then an infinite precision measurement is made on the rationals: the topology is discrete.
+If the value is measured with bounds, no infinite precision measurement is made and the value can be excluded but never fully verified: the topology is the one inherited from the reals.} It can be shown that a decidable domain, one where all statements are decidable, can only have up to countably many possibilities.
+Therefore, if the possibilities are uncountable, not all statements can be decidable.
+
+\section{Experimental relationships and domain combination}
+
+\partitle{Inference relationships} Intuitively, an \textbf{ inference relationship} between two domains $\edomain_X$ and $\edomain_Y$ allows us to infer truths on the first by performing experimental verification on the second.
+That is, an inference relationship is a map $\erel : \edomain_Y \to \edomain_X$ that given a verifiable statement in $\stmt_Y \in \edomain_Y$ gives us an equivalent statement $\stmt_Y \equiv \erel(\stmt_Y) \in \edomain_X$.\footnote{For example, by looking at the height of a column of mercury we are able to infer the temperature of the body.}
+
+\partitle{Causal relationship} Intuitively, a \textbf{causal relationship} between two domains $\edomain_X$ and $\edomain_Y$ means that if we determine which possibility $x \in X$ is true in the first domain, then this will also tell us what possibility $y \in Y$ is true in the second.
+That is, a causal relationship is a map $f : X \to Y$ such that $x \narrower f(x)$: if $x$ is true so is $f(x)$.
+An important result is that \textbf{causal relationships must be continuous} in the natural topology.
+Continuity, therefore, is a necessary requirement, not a desired property.
+We stress that this is topological continuity, and not necessarily analytical continuity.\footnote{Analytical continuity breaks down on discrete points, which mean on decidable possibilities.
+For example, whether a mixture of water, ice and water vapor is at the triple point of water is decidable, and therefore we can expect analytical discontinuities there, and in other phase transitions.}
+
+\partitle{Experimental relationships and domain dependence} Two domains admit an inference relationship if and only if they admit a causal relationship.
+We use the term \textbf{experimental relationship} to describe the link between the domains and we say $\edomain_X$ \textbf{depends on} $\edomain_Y$, noted $\edomain_X \subseteq \edomain_Y$.
+Note that the inference works in the opposite direction of the cause.
+The link between inference and causation is very important.
+Inference relationships are defined over verifiable statements, which are our link to experimental verification.
+This is desirable to have the most direct connection between relationships and experimental verification.
+However, inference our highly redundant\footnote{For example, an inference relationship will tell us how every finite precision measurement of the height of a mercury column would correspond to finite precision measurements of temperature.} and therefore they are less prominently used.
+On the other hand, causal relationships are defined over the possibilities, which are less directly related to experimental verification.
+Yet, they are a lot more compact\footnote{For example, a causal relationship will tell us what infinitely precise height of the mercury column will correspond to each infinitely precise value of temperature.} and therefore more practical.
+The link between them tells us that we can always go between one and the other, so we can use causal relationships almost exclusively.
+
+\partitle{Domain combination} Given a countable set of experimental domains $\{\edomain_X\}_{i=1}^{\infty}$, the combined domain is the experimental domain of all the verifiable statements generated by the other domain.
+Formally, it is the closure under finite conjunction and countable disjunction.
+The possibilities of the combined domain can be understood as a subset of a scalar product of the possibilities of all domains.
+The fewer correlations between the different domains, the bigger the set of combined possibilities.
+If the domains are independent, all combinations in the scalar product will correspond to a possibility for the combined domain.
+Moreover, the topology will be the product topology of the domains.
+If one domain depends on another, their combined domain will be equivalent to the second: no new possibilities are added.
+If the domains are incompatible, then the possibilities of the combined domain are the disjoint union and the topology will be the disjoint union topology.
+
+\partitle{Relationship domain} Given two experimental domain, we can construct a domain where the possibilities are the possible relationships between the domain.
+We call this the \textbf{relationship domain}.
+We can proceed recursively and construct domains of higher order functions, without going outside the mathematical framework.
+There is, however, a caveat: the construction of this domain requires the existence and construction of a logical meta-context.
+Therefore, while we can show that mathematically the construction can always be done, there is no guarantee that the statements are verifiable.
+That is: we cannot construct tests that verify relationships between quantities from tests that simply verify the values of the quantities.\footnote{That is, being able to verify particular values of temperature and pressure is not enough to be able to verify whether to a particular value of temperature always corresponds a particular value of pressure.}
 

### 2021-03-05 Overviews/Experimental verifiability/ExperimentalVerifiabilityOverview.tex
Another pass at exp ver overview

@@
 We briefly show how the use of topological spaces and $\sigma$-algebras in physics can be rederived and understood as the fundamental requirement of experimental verifiability.
-A set of experimentally distinguishable objects will necessarily be endowed with a topology that is Kolmogorov (i.e.
-$T_0$) and second countable, which puts severe constrains on the space of well-formed scientific theories.
-These idea can be taken as a first step in a general mathematical theory for experimental science.
+We will see that a set of experimentally distinguishable objects will necessarily be endowed with a topology that is Kolmogorov (i.e.
+$T_0$) and second countable, which both puts constraints on well-formed scientific theories and allow us to give concrete physical meaning to the mathematical constructs.
+These insights can be taken as a first step in a general mathematical theory for experimental science.
 This work is an overview of some of the results of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
@@
 
-The set of points of a topological space represents the possible cases (e.g.
-all possible values for the mass of a particle).
-Each element in the topology (i.e.
-each open set) represents a statement that can be experimentally verified (e.g.
-the mass of the particle is within a finite precision interval).
-Each element in the Borel algebra represent a statement that, though it may or may not be experimentally verified, is sound within the theory (e.g.
-the mass of the particle is exactly zero, the mass of the particle is an irrational number).
-The Assumptions of Physics framework rederives this structure from the idea that a scientific theory must be logically consistent and evidence based.
+A theoretical domain $\tdomain_X$ contains all the statements about a scientific object that can be associated to an experimental test, which we call theoretical statements.\footnote{For example, all the possible statements about the mass of a particle.} Within this set we find the possibilities which describe the possible cases that can be experimentally distinguished\footnote{For example, \statement{the mass of the particle is exactly zero.}}.
+Not all tests are guaranteed to always terminate,\footnote{For example, infinite precision measurements are only possible as the limit of an infinite sequence of finite precision measurements.} so we find the subset $\edomain_X$ of the verifiable statements, called experimental domain, for which successful termination coincides with the truth of the statement.\footnote{For example, \statement{the mass of the particle is less than 100 GeV.}}
+
+The possibilities corresponds to the points of our spaces.
+The experimental domain corresponds to the topology, where each open set corresponds to an open set.\footnote{For example, \statement{the mass of the particle is less than 100 GeV} corresponds to the open set $[0, 100)$.} The theoretical domain corresponds to the Borel algebra.
 
@@
 \partitle{Principle of scientific objectivity} The guiding principle is that \emph{science is universal, non-contradictory and evidence based.} This means that any scientific theory must specify a set of \textbf{statements} that are logically consistent, whose truth can be confirmed experimentally by everybody.
-This starting point is codified in the axioms that form the formal foundation of the theory.
+This starting point is codified in the following axioms, that form the formal foundation of our theory.
 
@@
 The \emph{axiom of context} tells us that statements are organized into a \textbf{logical context} $\logCtx$, and a well defined truth value for each statement must exist.
-This recognizes that the logical consistency and meaning of each statement relies on definitions and relationships that depends on the context\footnote{For example, if one is performing particle identification, the mass of an electron is assumed to be given.
+This recognizes that the logical consistency and meaning of each statement relies on definitions and relationships that depend on the context\footnote{For example, if one is performing particle identification, the mass of an electron is assumed to be given.
 If one is measuring the mass, the identify is assumed to be given, which the value of the mass is unknown.}, on a set of statements.
@@
 statements that must be equally true or false); narrowness (i.e.
-a statement is more specific than another)\footnote{For example, \statement{the mass of the object is between 4 and 5 kilograms} is narrower than \statement{the mass of the object is between 0 and 100 kilograms}.}; compatibility $\comp$ (i.e.
+a statement is more specific than another)\footnote{For example, \statement{the mass of the object is between 4 and 5 kilograms} is narrower than \statement{the mass of the object is between 0 and 100 kilograms}.
+Narrowness can also be understood as implication.}; compatibility $\comp$ (i.e.
 whether two statement can be true at the same time); independence $\indep$ (i.e.
 whether the truth of some statements constrains the truth of others).
-This basic structure constrains any further mathematical structure as the two will have to interact appropriately.
+This basic structure constrains any further mathematical structure as they will have to interact appropriately.
 
@@
 The \emph{axiom of verifiability} tells us that some statements are \textbf{verifiable}, meaning we have a test at our disposal that will terminate successfully in finite time if and only if the statement is true.
-Note that the test is not guaranteed to terminate, to give an answer, if the test is false.\footnote{For example, a test for \statement{there exists extra-terrestrial life} may terminate successfully if life is found, but it will not be conclusive if life is not found.
+Note that the test is not guaranteed to terminate, to give an answer, if the statement is false.\footnote{For example, a test for \statement{there exists extra-terrestrial life} may terminate successfully if life is found, but it will not be conclusive if life is not found.
 On the other hand, \statement{the mass of the particle is exactly 0 eV} can be experimentally disproved, but can never be confirmed due to finite precision.} Therefore the negation of a verifiable statement is not necessarily verifiable.
@@
 
-The remaining two axioms tell us that the finite conjunction and the countable disjunction of verifiable statements is verifiable.
-That is, we can use the tests of the statement to construct the test for the logical AND (i.e.
+The remaining two axioms tell us that the finite conjunction and the countable disjunction of verifiable statements are verifiable.
+That is, we can use the tests of the given statements to construct the test for the logical AND (i.e.
 check that all tests terminates successfully) and the logical OR (i.e.
-check that at least one test terminates).
+check that at least one test terminates successfully).
 The disjunction can be extended to a countable set because we only need one test to terminate successfully in a finite time, not all of them.
@@
 all the statements that describe the state of a system or the properties of an object).
-We call such a group an \textbf{experimental domain} $\edomain$ and it will have at least two characteristics.
+We call such a group an \textbf{experimental domain} $\edomain_X \subseteq \logCtx$ and it will have at least two characteristics.
 First of all, it will be closed under finite conjunction (i.e.
@@
 
-\partitle{Theoretical domains} While the verifiable statements define all the testing procedures that are available in some domain, they are not the only valid statements in a scientific theory.\footnote{For example, \statement{the mass of the photon is exactly 0 eV} is not something we can verify due to the limited precision of our statements, yet it is something we want to be able to say in our theory.} We are interested in all statement that can be assigned an experimental test, regardless of its terminating conditions.
-We call these \textbf{theoretical statements} and we group them into a \textbf{theoretical domain} $\tdomain$.
+\partitle{Theoretical domains} While the verifiable statements define all the testing procedures that are available in some domain, they are not the only valid statements in a scientific theory.\footnote{For example, \statement{the mass of the photon is exactly 0 eV} is not something we can verify due to the limited precision of our statements, yet it is something we want to be able to say in our theory.} We are interested in all statements that can be assigned an experimental test, regardless of its terminating conditions.
+We call these \textbf{theoretical statements} and we group them into a \textbf{theoretical domain} $\tdomain_X$.
 This is generated from experimental domain by closing under negation and countable disjunction.
@@
 
-\partitle{Correspondence to topologies and $\sigma$-algebras} Every statement in the domain can be characterize by the set of possibilities for which the statement is true.
-Therefore, we have a correspondence between sets and statement that preserves the structure.
+\partitle{Correspondence to topologies and $\sigma$-algebras} Every statement in the domain can be characterized by the set of possibilities for which the statement is true.
+Therefore, we have a correspondence between sets and statements where the logic operations become the set operations.
 Under this correspondence, the experimental domain induces a topology where each open set represents a verifiable statement, while the theoretical domain induces a $\sigma$-algebra where each Borel set is a theoretical statement.
@@
 
-\partitle{Experimental relationships and domain dependence} Two domains admit an inference relationship if and only if they admit a causal relationship.
+\partitle{Experimental relationships and domain dependence} An important result is that \textbf{two domains admit an inference relationship if and only if they admit a causal relationship}.
 We use the term \textbf{experimental relationship} to describe the link between the domains and we say $\edomain_X$ \textbf{depends on} $\edomain_Y$, noted $\edomain_X \subseteq \edomain_Y$.
 Note that the inference works in the opposite direction of the cause.
+
 The link between inference and causation is very important.
-Inference relationships are defined over verifiable statements, which are our link to experimental verification.
-This is desirable to have the most direct connection between relationships and experimental verification.
-However, inference our highly redundant\footnote{For example, an inference relationship will tell us how every finite precision measurement of the height of a mercury column would correspond to finite precision measurements of temperature.} and therefore they are less prominently used.
-On the other hand, causal relationships are defined over the possibilities, which are less directly related to experimental verification.
-Yet, they are a lot more compact\footnote{For example, a causal relationship will tell us what infinitely precise height of the mercury column will correspond to each infinitely precise value of temperature.} and therefore more practical.
-The link between them tells us that we can always go between one and the other, so we can use causal relationships almost exclusively.
+Inference relationships are defined over verifiable statements, which are our link to experimental verification, and therefore have a higher status in the theory.
+Unfortunately they are highly redundant\footnote{For example, an inference relationship will tell us how every finite precision measurement of the height of a mercury column would correspond to finite precision measurements of temperature.} and therefore harder to work with.
+On the other hand, causal relationships are defined over the possibilities, which are less directly related to experimental verification, but they are a lot more compact\footnote{For example, a causal relationship will tell us what infinitely precise height of the mercury column will correspond to each infinitely precise value of temperature.} and therefore more practical.
+The link between them tells us that we can always go between one and the other, so we can use causal relationships knowing that they are as experimentally well-defined as their counterparts.
 
-\partitle{Domain combination} Given a countable set of experimental domains $\{\edomain_X\}_{i=1}^{\infty}$, the combined domain is the experimental domain of all the verifiable statements generated by the other domain.
-Formally, it is the closure under finite conjunction and countable disjunction.
-The possibilities of the combined domain can be understood as a subset of a scalar product of the possibilities of all domains.
-The fewer correlations between the different domains, the bigger the set of combined possibilities.
+\partitle{Domain combination} Given a countable set of experimental domains $\{\edomain_X\}_{i=1}^{\infty}$, the combined domain is the experimental domain of all the verifiable statements generated by them.
+Formally, it is the closure under finite conjunction and countable disjunction.\footnote{For example, we take the domains of different quantities of an object to form the domain of its state.} The possibilities of the combined domain can be understood as a subset of a scalar product of the possibilities of all domains.\footnote{For example, the value of all the quantities provided by each domain.} The fewer correlations between the different domains, the bigger the set of combined possibilities.
 If the domains are independent, all combinations in the scalar product will correspond to a possibility for the combined domain.
@@
 
-\partitle{Relationship domain} Given two experimental domain, we can construct a domain where the possibilities are the possible relationships between the domain.
+\partitle{Relationship domain} Given two experimental domains, we can construct a domain where the possibilities are the possible relationships between the domain.
 We call this the \textbf{relationship domain}.

### 2021-03-10 Overviews/Experimental verifiability/ExperimentalVerifiabilityOverview.tex
Final pass at the experimental verifiability overview

@@
 We will see that a set of experimentally distinguishable objects will necessarily be endowed with a topology that is Kolmogorov (i.e.
-$T_0$) and second countable, which both puts constraints on well-formed scientific theories and allow us to give concrete physical meaning to the mathematical constructs.
+$T_0$) and second countable, which both puts constraints on well-formed scientific theories and allows us to give concrete physical meaning to the mathematical constructs.
 These insights can be taken as a first step in a general mathematical theory for experimental science.
@@
 A theoretical domain $\tdomain_X$ contains all the statements about a scientific object that can be associated to an experimental test, which we call theoretical statements.\footnote{For example, all the possible statements about the mass of a particle.} Within this set we find the possibilities which describe the possible cases that can be experimentally distinguished\footnote{For example, \statement{the mass of the particle is exactly zero.}}.
-Not all tests are guaranteed to always terminate,\footnote{For example, infinite precision measurements are only possible as the limit of an infinite sequence of finite precision measurements.} so we find the subset $\edomain_X$ of the verifiable statements, called experimental domain, for which successful termination coincides with the truth of the statement.\footnote{For example, \statement{the mass of the particle is less than 100 GeV.}}
+Not all tests are guaranteed to always terminate,\footnote{For example, infinite precision measurements are only possible as the limit of an infinite sequence of finite precision measurements.} so we find the subset $\edomain_X \subseteq \tdomain_X$ consisting of the verifiable statements, called experimental domain, for which successful termination coincides with the truth of the statement.\footnote{For example, \statement{the mass of the particle is less than 100 GeV.}}
 
-The possibilities corresponds to the points of our spaces.
-The experimental domain corresponds to the topology, where each open set corresponds to an open set.\footnote{For example, \statement{the mass of the particle is less than 100 GeV} corresponds to the open set $[0, 100)$.} The theoretical domain corresponds to the Borel algebra.
+The possibilities correspond to the points of our spaces.
+The experimental domain corresponds to the topology, where each open set corresponds to a verifiable statement.\footnote{For example, \statement{the mass of the particle is less than 100 GeV} corresponds to the open interval $(0, 100)$.} The theoretical domain corresponds to the Borel algebra.
 
@@
 
-\partitle{Principle of scientific objectivity} The guiding principle is that \emph{science is universal, non-contradictory and evidence based.} This means that any scientific theory must specify a set of \textbf{statements} that are logically consistent, whose truth can be confirmed experimentally by everybody.
-This starting point is codified in the following axioms, that form the formal foundation of our theory.
+\partitle{Principle of scientific objectivity} The guiding principle is that \emph{science is universal, non-contradictory and evidence based.} This means that any scientific theory must specify a set of \textbf{statements} that are logically consistent, whose truth is the same for everybody and connected to experimental verification.
+This starting point is codified in the following axioms, which form the formal foundation of our theory.
 
@@
 This recognizes that the logical consistency and meaning of each statement relies on definitions and relationships that depend on the context\footnote{For example, if one is performing particle identification, the mass of an electron is assumed to be given.
-If one is measuring the mass, the identify is assumed to be given, which the value of the mass is unknown.}, on a set of statements.
+If one is measuring the mass, the identity is assumed to be given, while the value of the mass is unknown.}, on a set of statements.
 
-The \emph{axiom of possibility} tell us that the statements in a context cannot be assigned arbitrary truth values, but only in a way that is consistent with the meaning defined by the context.\footnote{For example, \statement{the mass of the object is one kilogram} and \statement{the mass of the object is two kilograms} cannot be both true.} Therefore each context comes with a set $\pAss$ of possible assignments, where each possible assignment is a function $a : \logCtx \to \Bool$ that returns a value of truth for each statement.
-The truth is a possible assignment.
+The \emph{axiom of possibility} tell us that the statements in a context cannot be assigned arbitrary truth values, but only those consistent with the meaning defined by the context.\footnote{For example, \statement{the mass of the object is one kilogram} and \statement{the mass of the object is two kilograms} cannot both be true.} Therefore each context comes with a set $\pAss$ of possible assignments, where each possible assignment is a function $a : \logCtx \to \Bool$ that returns the truth value for each statement.
 
-The \emph{axiom of closure} tells us that we can always find a statements whose truth arbitrarily depends on the truth of other statements.
+The \emph{axiom of closure} tells us that we can always find a statement whose truth arbitrarily depends on the truth of other statements.
 
-\partitle{Logical relationships} The logic axioms entail that each logical context is a complete Boolean algebra.
-They allow us to define useful relationship such as: equivalence $\equiv$ (i.e.
+\partitle{Logical relationships} The axioms of logic entail that each logical context is a complete Boolean algebra.
+They allow us to define useful relationships such as: equivalence $\equiv$ (i.e.
 statements that must be equally true or false); narrowness (i.e.
@@
 whether two statement can be true at the same time); independence $\indep$ (i.e.
-whether the truth of some statements constrains the truth of others).
-This basic structure constrains any further mathematical structure as they will have to interact appropriately.
+whether the truth of some statements leaves the truth of others unconstrained).
+This basic structure constrains any further mathematical structure as the structures will have to interact appropriately.
 
@@
 
-The remaining two axioms tell us that the finite conjunction and the countable disjunction of verifiable statements are verifiable.
+The remaining two axioms state that the finite conjunction and the countable disjunction of verifiable statements are verifiable.
 That is, we can use the tests of the given statements to construct the test for the logical AND (i.e.
-check that all tests terminates successfully) and the logical OR (i.e.
+check that all tests terminate successfully) and the logical OR (i.e.
 check that at least one test terminates successfully).
@@
 
-\partitle{Decidability} A \textbf{falsifiable} statement is one whose negation is falsifiable.
+\partitle{Decidability} A \textbf{falsifiable} statement is one whose negation is verifiable.
 A \textbf{decidable} statement is one that is both verifiable and falsifiable.
@@
 
-\partitle{Different algebras} The logical context as a whole, the set of verifiable statements and the set of decidable statements follow different algebras according to table \ref{algebras}.
+\partitle{Different algebras} The logical context as a whole, the set of verifiable statements and the set of decidable statements follow different algebras according to Table \ref{algebras}.
 
@@
 Second, it cannot be too big: we must be able to explore it fully given an arbitrarily long amount of time.
-Formally, it can be generated from a countable subset of verifiable statements through finite conjunction and countable disjunction, it must have a countable \textbf{base}.
+Formally, it can be generated from a countable subset of verifiable statements through finite conjunction and countable disjunction, that is, it must have a countable \textbf{base}.
 This means the tests for all statements in the domain can be generated from the tests of the base, therefore running those countably many tests is enough to test the whole domain.
@@
 We call these \textbf{theoretical statements} and we group them into a \textbf{theoretical domain} $\tdomain_X$.
-This is generated from experimental domain by closing under negation and countable disjunction.
+This is generated from the experimental domain by closing under negation and countable disjunction.
 
@@
 These are statements that give the narrowest description possible and, if assumed to be true, tell us the truth value for all other statements.
-The set of possibilities represent all the possible cases that can be distinguished by the experimental domain.\footnote{For example, if the domain characterizes the state of a bead along the wire, ``the position of the bead is 3 meters and the velocity is zero meters per second'' is a possibility because it characterizes the truth value of all possible experimentally verifiable finite precision descriptions.}
+The set of possibilities represents all the possible cases that can be distinguished by the experimental domain.\footnote{For example, if the domain characterizes the state of a bead along a wire, ``the position of the bead is 3 meters and the velocity is zero meters per second'' is a possibility because it characterizes the truth value of all possible experimentally verifiable finite precision descriptions.}
 
@@
 Under this correspondence, the experimental domain induces a topology where each open set represents a verifiable statement, while the theoretical domain induces a $\sigma$-algebra where each Borel set is a theoretical statement.
-The topology is at least $T_0$ and second countable while the Borel algebra is countably generated (and therefore separable), which make them extremely well-behaved.
+The topology is at least $T_0$ and second countable while the Borel algebra is countably generated (and therefore separable), which makes them extremely well-behaved.
 
-\partitle{On the significance of mathematical structures} \textbf{The existence of well-behaved topologies and $\sigma$-algebra is a direct consequence of the requirement of experimental verifiabiliy.}\footnote{Philosophically, one may say that it is not an ontological attribute of the system being studied, but rather a characterization of its epistemically accessible characteristics.
-There is no reason to assume that ontological entities would follow such constrain.} Another important aspect is that \textbf{the points are not the starting point, but rather they are generated from the verifiable statements.} If we change what is being characterized (e.g.
+\partitle{On the significance of mathematical structures} \textbf{The existence of well-behaved topologies and $\sigma$-algebras is a direct consequence of the requirement of experimental verifiabiliy.}\footnote{Philosophically, one may say that it is not an ontological attribute of the system being studied, but rather a characterization of its epistemically accessible characteristics.
+There is no reason to assume that ontological entities would follow such constraint.} Another important aspect is that \textbf{the points are not the starting point, but rather they are generated from the verifiable statements.} If we change what is being characterized (e.g.
 add measurements of spin to those of the position and momentum of an electron) the points of the space will change.\footnote{Philosophically, the points themselves cannot be understood as some complete ontological entity, but rather the epistemically accessible characteristics under a set of given circumstances.}
 
-\partitle{Unphysical statements} Note that the theoretical domain is closed under countable operations, while the context is closed under arbitrary operations.
-This means that there may be statements that depend on the experimental ones that are not theoretical, they are not associated to a testing procedure: they are unphysical.
+\partitle{Unphysical statements} Our construction makes it clear that non-Borel sets play no role in physical theories, not because they are hard to work with, but because they are unphysical.
+Note that the theoretical domain is closed under countable operations, while the context is closed under arbitrary operations.
+This means that, in line of principle, there may be statements that depend on the experimental ones that are not theoretical.
 In fact, if the possibilities are uncountable (e.g.
 we use real numbers to identify them, as we do when using position for states) such statements are guaranteed to exist.
-These correspond to non-Borel sets, which therefore have no physical meaning.
+Since they are not theoretical, they are not associated to a testing procedure: they are unphysical and are automatically excluded by our construction.
+It is a feature of our theory that tells us exactly which mathematical objects are physically meaningful and which aren't.
 
-\partitle{Maximum cardinality for the possibilities.} Since the domain is generated by a countable set of verifiable statements, the possibilities must be distinguished with a countable sequence of boolean values.
+\partitle{Maximum cardinality for the possibilities} Since the domain is generated by a countable set of verifiable statements, the possibilities must be distinguished with a countable sequence of Boolean values.
 This means that the cardinality of the possibilities cannot exceed that of the continuum (an infinite sequence of zeros and ones is a real number expressed in base 2), therefore infinities of higher order and all associated problems do not play a role in science.
@@
 the set of possibilities for which its test terminates successfully), its falsifiable part (i.e.
-test terminates successfully) and the undecidable part (i.e.
+the test terminates successfully) and the undecidable part (i.e.
 the test does not terminate).
-Topologically, these correspond to the interior, exterior and boundary of the corresponding Borel set.
-\footnote{Note there are domains that contain undecidable statements, those for which the test never terminates.
-For example, \statement{the mass of the proton is rational in units of eV} is undecidable: any finite precision measurement will be consistent with infinitely many rationals and irrational possibilities.
-Mathematically, the boundary of the rational is all the reals.}
+Topologically, these correspond to the interior, exterior and boundary of the corresponding Borel set.\footnote{Note there are domains that contain undecidable statements, those for which the test never terminates.
+For example, \statement{the mass of the proton is rational in units of eV} is undecidable: any finite precision measurement will be consistent with infinitely many rational and irrational possibilities.
+Mathematically, the boundary of the rationals is all the reals.}
 
 \partitle{Discreteness is decidability} A decidable statement will correspond to a clopen set.
-A topology is discrete if all sets are open, which means all the statements are decidable.\footnote{Many physicists confuse discreteness with countability, and characterize the rationals as discrete.
+A topology is discrete if all sets are clopen, which means all the statements are decidable.\footnote{Many physicists confuse discreteness with countability, and characterize the rationals as discrete.
 This depends on how the rationals are measured.
@@
 \partitle{Inference relationships} Intuitively, an \textbf{ inference relationship} between two domains $\edomain_X$ and $\edomain_Y$ allows us to infer truths on the first by performing experimental verification on the second.
-That is, an inference relationship is a map $\erel : \edomain_Y \to \edomain_X$ that given a verifiable statement in $\stmt_Y \in \edomain_Y$ gives us an equivalent statement $\stmt_Y \equiv \erel(\stmt_Y) \in \edomain_X$.\footnote{For example, by looking at the height of a column of mercury we are able to infer the temperature of the body.}
+That is, an inference relationship is a map $\erel : \edomain_Y \to \edomain_X$ that given a verifiable statement $\stmt_Y \in \edomain_Y$ gives us an equivalent statement $\stmt_Y \equiv \erel(\stmt_Y) \in \edomain_X$.\footnote{For example, by looking at the height of a column of mercury we are able to infer the temperature of the body.}
 
@@
 Continuity, therefore, is a necessary requirement, not a desired property.
-We stress that this is topological continuity, and not necessarily analytical continuity.\footnote{Analytical continuity breaks down on discrete points, which mean on decidable possibilities.
+We stress that this is topological continuity, and not necessarily analytical continuity.\footnote{Analytical continuity breaks down on discrete points, which means on decidable possibilities.
 For example, whether a mixture of water, ice and water vapor is at the triple point of water is decidable, and therefore we can expect analytical discontinuities there, and in other phase transitions.}
@@
 Inference relationships are defined over verifiable statements, which are our link to experimental verification, and therefore have a higher status in the theory.
-Unfortunately they are highly redundant\footnote{For example, an inference relationship will tell us how every finite precision measurement of the height of a mercury column would correspond to finite precision measurements of temperature.} and therefore harder to work with.
-On the other hand, causal relationships are defined over the possibilities, which are less directly related to experimental verification, but they are a lot more compact\footnote{For example, a causal relationship will tell us what infinitely precise height of the mercury column will correspond to each infinitely precise value of temperature.} and therefore more practical.
+Unfortunately they are highly redundant\footnote{For example, an inference relationship will tell us how every finite precision measurement of the height of a mercury column would correspond to a finite precision measurement of temperature.
+As the finite intervals overlap in infinitely many ways, the overlaps are specified again and again.} and therefore harder to work with.
+On the other hand, causal relationships are defined over the possibilities, which are less directly related to experimental verification, but they are a lot more compact\footnote{For example, a causal relationship will tell us what infinitely precise height of the mercury column will correspond to each infinitely precise value of temperature.
+Each value is mapped once to another value.} and therefore more practical mathematically.
 The link between them tells us that we can always go between one and the other, so we can use causal relationships knowing that they are as experimentally well-defined as their counterparts.
@@
 \partitle{Domain combination} Given a countable set of experimental domains $\{\edomain_X\}_{i=1}^{\infty}$, the combined domain is the experimental domain of all the verifiable statements generated by them.
-Formally, it is the closure under finite conjunction and countable disjunction.\footnote{For example, we take the domains of different quantities of an object to form the domain of its state.} The possibilities of the combined domain can be understood as a subset of a scalar product of the possibilities of all domains.\footnote{For example, the value of all the quantities provided by each domain.} The fewer correlations between the different domains, the bigger the set of combined possibilities.
-If the domains are independent, all combinations in the scalar product will correspond to a possibility for the combined domain.
+Formally, it is the closure under finite conjunction and countable disjunction.\footnote{For example, we take the domains of different quantities of an object to form the domain of its state.} The possibilities of the combined domain can be understood as a subset of the Cartesian product of the possibilities of all domains.\footnote{For example, the value of all the quantities provided by each domain.} The fewer correlations between the different domains, the bigger the set of combined possibilities.
+If the domains are independent, all combinations in the Cartesian product will correspond to a possibility for the combined domain.
 Moreover, the topology will be the product topology of the domains.
@@
 
-\partitle{Relationship domain} Given two experimental domains, we can construct a domain where the possibilities are the possible relationships between the domain.
+\partitle{Relationship domain} Given two experimental domains, we can construct a domain where the possibilities are the possible relationships between the domains.
 We call this the \textbf{relationship domain}.
@@
 Therefore, while we can show that mathematically the construction can always be done, there is no guarantee that the statements are verifiable.
-That is: we cannot construct tests that verify relationships between quantities from tests that simply verify the values of the quantities.\footnote{That is, being able to verify particular values of temperature and pressure is not enough to be able to verify whether to a particular value of temperature always corresponds a particular value of pressure.}
+That is, we cannot construct tests that verify relationships between quantities from tests that simply verify the values of the quantities.\footnote{That is, being able to verify particular values of temperature and pressure is not enough to be able to verify whether to a particular value of temperature always corresponds a particular value of pressure.}
 

### 2021-03-14 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Updates to various parts

@@
 
-If a system is infinitesimally reducible, then we can define a distribution $\rho$ over particle states.
-Furthermore, since particle states are independent of a choice of units, this distribution must be invariant.
-We show in Section~\ref{infinitesimal} that defining this invariant distribution requires a symplectic manifold, which serves as the phase space of classical mechanics.
+If a system is infinitesimally reducible, then we can define a distribution $\rho$ over particle states and a density $\rho_{\xi}$ over the corresponding state variables $\xi$.
+Since particle states are independent of a choice of units, this density must be invariant.
+We show in Section~\ref{infinitesimal} that defining this invariant density requires a symplectic manifold, which serves as the phase space of classical mechanics.
 The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
@@
 Secondly, a system has \textit{deterministic and reversible evolution} provided that the present state of the system uniquely determines both its future and past states.
-This condition entails that the invariant distribution $\rho$ is in fact a conserved phase space density.
+This condition entails that the invariant density $\rho_{\xi}$ is a conserved phase space density.
 We will show that a classical system which satisfies this condition is Hamiltonian: its phase space is a symplectic manifold that evolves according to Hamiltonian evolution  (i.e.
@@
 Finally, our treatment is not restricted to $n$-particle mechanics but applies as well to rigid-body and continuum mechanics.\footnote{For discussion of the numerous conceptual problems that arise within different approaches to classical mechanics (including point-masses, rigid-bodies, and continua) see \textcites{Wilson}.
-In Wilson's terminology, the framework we apply here may be yet another `theory facade' for classical physics, but a fruitful one nonetheless.} Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
+In Wilson's terminology, the framework we apply here may be yet another `theory facade' for classical physics, but a fruitful one nonetheless.} This supports our contention that the property of infinitesimal reducibility fruitfully characterizes classical systems.
+Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
+%The property of infinitesimal reducibility provides a fruitful notion of a classical system.
+Classical systems are infinitesimally reducible systems.
+Hence, we feel this is the proper notion of a classical system.
 % For fluid systems, the deterministic/reversible evolution condition applies to the whole system but not for the individual particles due to interactions between these.}
@@
 
-Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant distributions (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus.
+Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant densities (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus.
 Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration.
@@
 Here, $\xi(s) \equiv \xi^a$ and $\hat{\xi}(s) \equiv \hat{\xi}^b$.} Likewise, the integral $\int_U \rho_c(s) d\mathcal{S} $ is independent of the choice of state variables, so $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.
-By the change of variables formula, we also know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\phi = \hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
+By the change of variables formula, we also know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\phi \equiv \hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
 For convenience, we denote $\rho_{c, \hat{\xi}} (\phi(\xi^a))$ by $\rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
@@
 
-As a scalar, the units of the distribution $\rho_c$ are invariant under changing the units of the state variables.
+As a scalar, the units of the density $\rho_c$ are invariant under changing the units of the state variables.
 So suppose we change the units by transforming $q$ to a new unit variable $\hat{q}=\hat{q}(q)$.
@@
 
-The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is a coordinate-invariant distribution over the state space $\mathcal{S}$ of its infinitesimal parts.
+The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is represented by a coordinate-invariant density over the state space $\mathcal{S}$ of its infinitesimal parts.
 We have shown that the state space of the particles must be a symplectic manifold.
-Additionally, any symplectic manifold structure supports the definition of coordinate-invariant distributions: any scalar function can be integrated using the symplectic form in a coordinate-invariant way.
+Additionally, any symplectic manifold structure supports the definition of coordinate-invariant densitys: any scalar function can be integrated using the symplectic form in a coordinate-invariant way.
 Therefore the symplectic structure is exactly the structure needed to define an invariant density.
@@
 Likewise, \textcites[164,171]{Wallace} claims that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
-Yet, as we now show, the conjugate momentum $p$ is actually gauge-dependent.
+Yet, as we now show, the conjugate momentum $p$ is actually gauge-dependent, even in the context of a free particle.
 Assuming that gauge-invariance is a necessary condition for fundamental physical significance, this precludes $p$ from having fundamental status.
@@
 In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
-Unlike kinetic momentum, conjugate momentum is not an experimentally measurable quantity, because its specification involves a gauge choice.
+Unlike both velocity and the kinetic momentum $m g_{ji} v^i$, conjugate momentum is not an experimentally measurable quantity, because its specification involves a gauge choice.
 
@@
 By analyzing classical systems in terms of their invariant phase space density, we gain new insights into Lagrangian mechanics.
-Section~\ref{action} shows how the action principle can be understood geometrically in the extended phase space as a consequence of ``state density" conservation.
+Section~\ref{action} shows how the action principle can be understood geometrically in the extended phase space as a consequence of the conserved density.
 We show that the action principle holds even when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
@@
 Although central to Lagrangian mechanics, we can actually derive the stationary action principle using Hamiltonian mechanics.
-Our derivation simply assumes that 1) the system trajectory follows a deterministic and reversible evolution and 2) the system satisfies the motion equivalence condition.
+Our derivation simply assumes that the system trajectory follows a deterministic and reversible evolution.
 We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a related argument, see \textcites[236--37, 243--45]{Arnold}.}
@@
 Formally, the vector field $\vec{S} := (\frac{d q }{d t }, \frac{d p }{d t }, \frac{d t }{d t })$ represents the displacement of states along the trajectories.
-This field is divergenceless, $\nabla \cdot \vector{S} = 0$, because given a region $U$ as many states will enter the region as exit.
+This field is divergenceless, $\nabla \cdot \vector{S} = 0$, because for any region $U$, the same number of states will enter the region as exit.
 In particular, note that the displacement is tangent to the trajectory $\gamma [q, p, t] (t)$ in extended phase space, where the variable $t$ functions as both a coordinate and the evolution parameter.
@@
 
-At this point, we return to the action $ \mathscr{A}$, defined as the integral of the Lagrangian over a trajectory beginning at time $t_0$ and ending at time $t_1$: $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt$.
-Provided that the Legendre transform is well-defined (i.e.
+At this point, we return to the action $ \mathscr{A}$.
+If the motion equivalence condition holds, then $ \mathscr{A}$ is defined as the integral of the Lagrangian over a trajectory beginning at time $t_0$ and ending at time $t_1$: $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt$.
+If motion equivalence does not hold, we can define the action more generally over the function $L_p \equiv p \dot{q} - H(q, p, t)$, considering paths in the extended phase space.
+By deriving the stationary action principle in this more general setting, we simultaneously derive its usual version within Lagrangian mechanics.\footnote{When motion equivalence holds, the Legendre transform is well-defined, allowing us to express $L_p$ as a function of $(q, \dot{q}, t)$, i.e.
+$L_p = L(q, \dot{q}, t) = p \dot{q} - H$.} We have then that $ \mathscr{A}[\gamma] = \integral^{t_1}_{t_0} L_p (q, p, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$, where $\gamma$ is an actual trajectory followed by the system.
+Multiplying the Hamiltonian by $dt/dt = 1$ and adding a vanishing $0 \cdot \frac{dp}{dt}$, we obtain $ \mathscr{A}[\gamma] = \integral^{t_1}_{t_0} [p \frac{dq}{dt} + 0 \cdot \frac{dp}{dt} - H(q, p, t) \frac{dt}{dt} ]dt$.
+We notice that the integrand simply equals the inner product of $\vector{\theta}$ with $\vector{S} $, entailing that $L_p =\vector{\theta} \cdot \vector{S} $.
+Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$.
+Note that $\vector{S}  dt = (\frac{d q }{d t } dt, \frac{d p }{d t } dt, \frac{d t }{d t }dt ) = (dq, dp, dt) = d\vector{\gamma}$ represents the infinitesimal displacement along the trajectory.
+Therefore, the action equals the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.
+To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
+  %Provided that the Legendre transform is well-defined (i.e.
 the motion equivalence assumption is satisfied), then $H = p \dot{q} - L$.
 Hence, we can re-express the action as $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$.
-Multiplying the Hamiltonian by $dt/dt = 1$ and adding a vanishing $0 \cdot \frac{dp}{dt}$, we obtain $ \mathscr{A} = \integral^{t_1}_{t_0} [p \frac{dq}{dt} + 0 \cdot \frac{dp}{dt} - H(q, p, t) \frac{dt}{dt} ]dt$.
-We notice that the integrand simply equals the inner product of $\vector{\theta}$ with $\vector{S} $, entailing that the Lagrangian $L =\vector{\theta} \cdot \vector{S} $.
-Consequently, the action $ \mathscr{A} [\gamma] = \integral^{t_1}_{t_0} \vector{\theta} \cdot \vector{S}  dt$ where $\gamma$ is an actual trajectory followed by the system.
-Note that $\vector{S}  dt = (\frac{d q }{d t } dt, \frac{d p }{d t } dt, \frac{d t }{d t }dt ) = (dq, dp, dt) = d\vector{\gamma}$ represents the infinitesimal displacement along the trajectory.
-Therefore, the action for those trajectories can be rewritten as the line integral of $\vector{\theta}$ along the trajectory $\gamma $: $ \mathscr{A} [\gamma] =\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$.
-To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
 % Note on the generality of the argument: in cases where motion equivalence is not satisfied, we can still write the integrand in terms of H(q, p, t), although it would not be an integral over the Lagrangian.
@@
 Recalling now that i) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and ii)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $curl(\vector{\theta})$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
-This establishes Hamilton's principle of stationary action, which we have derived by assuming that 1) the displacement vector is divergenceless and 2) using the Legendre transform (well-defined when motion equivalence holds).
+This establishes a more general Hamilton's principle of stationary action, which we have derived simply by assuming that the displacement vector is divergenceless.
+In the special case that motion equivalence holds, the Legendre transformation is well-defined and $L_p = p \dot{q} - H(q, p, t)$ equals the usual Lagrangian $L(q, \dot{q}, t)$, thereby recovering the usual principle of stationary action for Lagrangian mechanics.
 
@@
 As discussed in Section~\ref{fundamental}, the conjugate momentum $p$ is a gauge-dependent variable.
-In contrast, position and kinetic momentum are both gauge-invariant, making them physically privileged.
+In contrast, position and velocity are both gauge-invariant, making them physically privileged.
 Consequently, Lagrangian mechanics employs state space variables that are gauge-invariant, while Hamiltonian mechanics uses a gauge-dependent variable.
@@
 There, we saw how both the linear relationship between momentum and velocity and the quadratic form (\ref{6.2}) follow from imposing the motion equivalence condition.
-In Equation~\ref{linear_relationship}, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (connected to the gauge-dependency of the conjugate momentum and the electromagnetic field).
+In Equation~\ref{linear_relationship}, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (connected to the gauge-dependency of the conjugate momentum).
 Just as Curiel does, we then used this relationship to derive a quadratic form for the Hamiltonian (Equation~\ref{massive_Hamiltonian}).
@@
 These include classical models of a massless particle moving at constant speed, such as a classical photon.
-In general, Hamiltonian systems outside the hyper-regular domain correspond to cases where we cannot express the density of states in terms of position and velocity.
+In general, Hamiltonian systems outside the hyper-regular domain correspond to cases where we cannot express the density in terms of position and velocity.
 These are cases where the weak equivalence condition (\ref{weak_equivalence}) fails.
@@
 Of course, if one insists on defining a `classical system' as Curiel does, one has grounds for privileging Lagrangian mechanics.
-Nevertheless, our definition of classical systems has substantive advantages over Curiel's.
+Nevertheless, our preferred notion of classical systems---focusing on systems that are infinitesimally reducible---has substantive advantages over Curiel's.
 Preceding sections have illustrated many of these advantages.
-Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{density}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a more geometrical understanding of the principle of stationary action (Section~\ref{action}).
+Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{density}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a more geometrical and general understanding of the principle of stationary action (Section~\ref{action}).
 Additionally, our definition accommodates models of classical massless particles by maintaining the uniqueness of solutions, even outside the hyper-regular domain.
@@
 [omit for review].} By changing the requirement of infinitesimal reducibility (Section~\ref{infinitesimal}) to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.
-Hence, our definition perspicuously distinguishes quantum from classical systems.
+Hence, our framework perspicuously distinguishes quantum from classical systems.
 We take these considerations as reason enough for adopting our preferred notion of a classical system.
@@
 We have also argued that non-fundamental quantities can be physically significant despite being neither coordinate- nor gauge-invariant.
-Finally, we have provided a geometrical interpretation of the principle of stationary action that makes the gauge-dependency of the Lagrangian manifest.
+Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain while making the gauge-dependency of the Lagrangian manifest.
 Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but classical systems in general.

### 2021-03-18 Articles/2020-QuantumLogic/Source/quantumlogic.tex
First draft from Andrea

@@
-\documentclass[10pt,twocolumn, nofootinbib]{revtex4-1}
-%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
-%\documentclass[10pt,twocolumn,letterpaper]{article}
-
-\usepackage{amsmath}
+\documentclass[11pt, executivepaper]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{natbib}
+\usepackage{amsmath, amsthm, amssymb}
 \usepackage{amsfonts}
-\usepackage{amsthm}
 \usepackage{MnSymbol}
+\usepackage{mathtools}
+\usepackage{xcolor}
 \usepackage{graphicx}
 \usepackage{enumitem}
+\usepackage{geometry}
+ \geometry{
+ a4paper,
+ total={155mm,237mm},
+ left=28mm,
+ top=32mm,
+ }
 \usepackage{hyperref}
-\hypersetup{
-	colorlinks=true,
-	citecolor=blue,
-	urlcolor=blue,
-	linkcolor=blue
-}
-\urlstyle{same}
-\frenchspacing
+\hypersetup{colorlinks= true, allcolors=blue}
+\setcitestyle{aysep={}}
 
-\newtheorem{prop}[equation]{Proposition}
+\newtheorem{prop}{Proposition}
 
@@
 
-\title{Notes on classical/quantum logic }
-\author{Gabriele Carcassi, Christine A. Aidala}
-\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\title{\textbf{On the Common Logical Structure of Classical and Quantum Mechanics}}
 
-\date{\today}
-
-\begin{abstract}
-Notes of classical logic in quantum mechanics.
-\end{abstract}
+\author{Gabriele Carcassi\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109.
+E-mail: carcassi@umich.edu} \and Christine A.
+Aidala\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109.
+E-mail: caidala@umich.edu} \and Andrea Oldofredi\thanks{Section de Philosophie, Universit\'e de Lausanne, Switzerland.
+E-mail: Andrea.Oldofredi@unil.ch}}
 
@@
 
-\section{Quantum logic: definition and motivation}
+\begin{abstract}
+\center Notes on classical and quantum logic
+\vspace{4mm}
+
+\noindent \emph{Keywords}: Quantum Mechanics; Quantum Logic; Classical Logic; 
+\end{abstract}
+\vspace{5mm}
+\clearpage
+
+\tableofcontents
+\vspace{5mm}
+
+\section{Introduction}
+
+As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus, a belief that is widespread still to this day (cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004} and \cite{deRonde:2016}).
+From the thirties, in fact, various modifications of Classical Logic (CL) were advanced in order to faithfully represent the physical content of quantum theory.
+For instance, in 1931 the Polish philosopher Zygmunt Zawirski proposed to apply \L ukasiewicz's three-valued logic to QM, starting from considerations about the wave-particle duality and Heisenberg's uncertainty principle.
+In virtue of the latter, moreover, the logical empiricists Moritz Schlick and Philipp Frank claimed that the conjunction of two or more statements may have no meaning in quantum theory, concluding that the classical conjunction is not always valid in the context of quantum physics, as recalled in \cite{Carnap:1966}.
+In their view, a typical example of invalid conjunction is one attributing a definite value for both position and momentum to a single quantum particle at a given time $t$.
+Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it s generally recognized that
+the founding text on Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 von Neumann had already stated that from the algebraic structure of quantum theory it is possible to formulate a new propositional calculus, anticipating seminal ideas that would have been fully expressed of his successive work in collaboration with Garrett Birkhoff.}
+
+A few years after the publication of this paper, the research in quantum logic had a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct formal structures to represent the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}).
+In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography.
+Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now widespread among experts as exemplified e.g.\ by \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
+Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory].
+This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).}, it is an important discipline of contemporary physics.
+In this paper we will neither take a stance in the dispute concerning the ontological status of logic and its laws, nor we will consider recent developments of QL given that they are not strictly relevant to the point we are going to make.
+Indeed, the present essay is focused on the logical structures common to both quantum and Classical Mechanics (CM), following the spirit of the former discussion.\footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM.
+First order and higher order logics will not be discussed here as well as the logic of relativistic quantum theories.}
+
+Recently one of us asked in another essay whether quantum physics necessarily implies a revision of classical logic (cf.\ \cite{Oldofredi:2020}).
+Alternatively stated, it was asked whether classical logic must be abandoned in the quantum realm or there are quantum theories compatible with a classical propositional calculus.
+Analyzing Bohmian Mechanics, a well-known interpretation of QM, it has been showed that quantum physics does not necessarily involves a departure from classical logic.
+More precisely, it has been argued that in the context of BM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow CL---but also the distributivity law holds, contrary to the case of standard QM.
+This fact is a consequence of the clear ontology of the theory and its its theory of measurement.
+
+In this paper we will be concerned with a similar issue, since we ask whether standard quantum mechanics necessarily involves a new logical structure radically different with respect to classical propositional calculus.
+In what follows we will answer this question in the negative.
+More precisely, here we have a three-fold aim: in the first place, we will analyze the main motivations and justifications for the introduction of quantum logic, arguing that they do not entail the failure of CL in the quantum context.
+In the second place we argue that classical and quantum mechanics share a common algebraic structure which can originate a classical propositional calculus for quantum propositions.
+A position dissimilar to the received view about QL.
+Here we will also show that several important statements in quantum mechanics, i.e.\ those concerned with expectation values, are not captured by QL, while they may be taken into account in a classical logical system.
+Finally, we will argue that it is possible to show that classical propositions may obey a quantum logic.
+Clearly, we will also spell out the reasons for which CL seems not to work in the context of QM.
+All these aspects, to our knowledge, are not sufficiently discussed in literature, thus, the present paper can contribute to shed new light on the relation between logic and quantum mechanics.
+\vspace{2mm}
+
+The essay is structured as follows: Section \ref{QL} provides a brief outline of the standard approach to QL (readers familiar with Birkhoff and von Neumann's proposal can skip this part), while in Section \ref{Motivations} we review the main motivations that have been given in literature to introduce a quantum logical calculus, explaining also why they do not entail the failure of classical logic in QM.
+In Section \ref{Math} we provide a mathematical argument aimed at showing not only that QM is compatible with classical logic---sharing a common logical structure with classical mechanics---but also that important sentences of quantum theory are excluded by standard QL, as stated a few lines above.\ In Section \ref{Discussion} we discuss the main philosophical implications of the preceding sections, and Section \ref{conc} concludes the paper.
+
+\section{The Standard Approach to Quantum Logic}
+\label{QL}
+
+In order to present the standard approach to QL, let us say a few words about the relation between classical logic and classical mechanics.
+It is a well known fact that classical propositional logic is equivalent to a Boolean algebra where propositions (atomic and complex) have truth values ``true'' and ``false'' (or 1 and 0), and the main operations among them are conjunction ``$\wedge$'', disjunction ``$\vee$'' and negation ``$\neg$'', the only unary operation.
+From these logical connectives one can introduce secondary operations, as for instance the material implication ``$\longrightarrow$'', the exclusive or ``$\oplus$'' and logical equivalence ``$\longleftrightarrow$''.
+Furthermore, it is important for our discussion to underline that the laws of propositional logic include commutativity, associativity, identity and distributivity with respect to $\wedge$ and $\vee$.
+
+Analogously, the algebraic structure underlying classical mechanics is Boolean as well, being commutative, distributive and associative; in this context the logical operations of conjunction, disjunction and negation are replaced respectively by multiplication, addition and complementation among the variables of the theory.
+This fact has an interesting physical significance, since the variables associated to magnitudes of physical systems can be added and multiplied together---i.e.\ one can sum and multiply measurement results.
+Both propositional logic and classical mechanics, thus, generate a complemented Boolean lattice; from this fact it follows that the observable algebra of CM ``is isomorphic to a Boolean algebra of propositions with $(\wedge, \vee, \neg)$'' (\cite{David:2015}, p.
+79).
+To this regard, Beltrametti claims that 
+\begin{quote}
+Boolean algebras are algebraic models of classical logic (more specifically of classical propositional calculus) with the algebraic operations of meet ($\wedge$) and join ($\vee$) corresponding to the logical connectives ``and'', ``or'', and the unary relation of orthocomplementation corresponding to the logical negation.
+The rules and tautologies of classical logic have their counterpart in equations and identities of Boolean algebras (\cite{Beltrametti:2004}, p.\ 341).
+\end{quote}
+
+Therefore, it is possible to formally characterize the state of a physical system via logical propositions which can assume the truth values ``true'' or ``false''---depending whether such statements describe true or false state of affairs concerning the system under consideration (cf.\ \cite{Jaeger:2009}, p.
+61)---and to perform logical operations among propositions via the connectives $(\wedge, \vee, \neg)$.
+It is worth noting that in the context of CM it is always determined whether a system instantiates a given property or not; alternatively stated, every logical proposition about physical systems is either true or false, and hence in CM the principle of semantic bivalence holds (for more details cf.\ \cite{Bub:2007}, p.\ 642, \cite{Giuntini:2002}, p.\ 130).
+
+Taking instead into account the mathematical structure and the physical content of standard QM, it is generally claimed that the classical propositional calculus is no longer appropriate to represent the logic obeyed by quantum propositions, since quantum theory does not share the same algebraic structure of classical mechanics.
+In particular, the algebra of quantum observables is non-commutative and, from a logical perspective, generates a non-distributive orthocomplemented lattice, which is non-Boolean.
+Referring to this, given the axioms of QM Birkhoff and von Neumann showed how the physical content of the theory does not conform to a classical logical structure.
+The principal aim of their founding essay of quantum logic was to ``discover what logical structures one may hope to find in physical theories which, like quantum mechanics, do not conform to classical logic'' (\cite{vonNeumann:1936}, p.\ 823).
+The expression quantum logic in this paper must be clarified, since it refers to a quantum propositional calculus in the form of a ``calculus of linear subspaces with respect to \emph{set products}, \emph{linear sums}, and \emph{orthogonal complements}'' which ``resembles the usual calculus of propositions with respect to \emph{and}, \emph{or}, and \emph{not}'' (\emph{ibid}.), where the logical propositions are associated to measurements, tests on quantum systems.\footnote{Cf.\ \cite{Giuntini:2002} and \cite{DallaChiara:2004} for a systematic introduction to various forms of quantum logics, and to \cite{Engesser:2009} for historical and philosophical discussions on the topic.}
+
+Their analysis begins by defining quantum logical propositions as \emph{experimental propositions}, i.e.\ statements affirming that a certain observable $A$ (or a set of observables) measured on a quantum system $s$ has a given value $a_i$ (or a sequence of values).
+More precisely, the authors stress that both in classical and quantum mechanics observations of physical systems are given by readings of experimental outcomes ($x_1, \dots, x_n$) of \emph{compatible} measurements ($\mu_1, \dots, \mu_n$).
+The values ($x_1, \dots, x_n$) are elements of what the authors called the ($x_1, \dots, x_n$)-space, i.e.\ the ``observation-space'' of the system in question, whose elements are all the possible combinations of results of the compatible measurements ($\mu_1, \dots, \mu_n$).
+Hence, the actual values ($x_1, \dots, x_n$) form a subset of such a space.
+Birkhoff and von Neumann, then, defined the experimental propositions concerning a physical system as the subsets of the observation space associated with it.
+
+Secondly, the authors underlined that in classical and quantum mechanics the states of physical systems are mathematically represented by points in their state spaces---phase space for the classical case, and Hilbert space $\mathcal{H}$ for the quantum case---which provide the maximal information concerning the system.
+A point in phase space corresponds to the specification of the position and momentum variables of a certain classical system, whereas in QM the points of $\mathcal{H}$ correspond to wave functions.
+The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcome in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.} as underlined by Svozi:
+\begin{quote}
+Any closed linear subspace of---or, equivalently, any projection operator on---a Hilbert space corresponds to an elementary
+proposition. The elementary \emph{true--false} proposition can in English be spelled out explicitly as
+\begin{quote}
+``The physical system has a property corresponding to the associated closed linear subspace'' (\cite{Svozi:1999}, p.\ 1).
+\end{quote}
+\end{quote}
+
+Thus, since in quantum mechanics physical systems do not have well-defined values for their properties before the measurement of a given operator (contrary to their classical counterparts), or better that quantum observations do not reveal pre-existing values, we have to stress that such quantum propositions refer always to measurements or to preparation procedures.
+Hence, the sentence ``the particle $s$ has $x$-spin up'' means either that I have measured the observable $S_x$ and I have found the value $+\hbar/2$ associated with the state ``$x$-spin-up'', or equivalently that I prepared $s$ in the $x$-spin-up state.
+Thus, one cannot simply say, as in classical mechanics, that a certain system has a certain property without having prepared such a system in a given state or having measured it.\footnote{More on this below.}
+
+Now that we have clarified what quantum propositions are, let us introduce Birkhoff and von Neumann's quantum logic.
+In order to properly define a propositional calculus for QM, one must define the logical operators for conjunction, disjunction and negation and the notion of logical implication.
+Following \cite{vonNeumann:1936}, the procedure is rather simple:
+\begin{enumerate}
+\item The negation of a proposition $p$ is defined by the authors as follows: ``since all operators of quantum mechanics are Hermitian, the mathematical representative of the negative of any experimental proposition is the orthogonal complement of the mathematical representative of the proposition itself'' (\cite{vonNeumann:1936}, pp.\ 826-827).
+The orthogonal complement $P^{\perp}$ of the subspace $P$ is the set whose elements are all the vectors orthogonal to the elements of $P$.
+Such an orthogonal complement satisfies the following property: given a subset $P\subset\mathcal{H}$ and a pure state $\psi$, $\psi(P)=1$ iff $\psi(P^{\perp})=0$ and $\psi(P)=0$ iff $\psi(P^{\perp})=1$.
+As Dalla Chiara and Giuntini underline, ``$\psi$ assigns to an event probability 1 (0, respectively) iff $\psi$ assigns to the orthocomplement of $P$ [notation adapted] probability 0 (1, respectively).
+As a consequence, one is dealing with an operation that \emph{inverts} the two extreme probability-values, which naturally correspond to the truth-values \emph{truth} and \emph{falsity} (similarly to the classical truth-table of negation)'' (\cite{Giuntini:2002}, p.\ 132).
+\item Concerning the conjunction, Birkhoff and von Neumann notice that one can retain the very same set-theoretical interpretation of the classical conjunction in QL, since the intersection of two closed subspaces $P, Q\subset\mathcal{H}$ is still a closed subspace.
+Thus, one maintains the usual meaning for the conjunction: the pure state $\psi$ verifies $P\cap Q$ iff $\psi$ verifies both $P$ and $Q$.
+Thus, quantum logic does not introduce a new logical operator for the conjunction.
+\item Contrary to the previous case, the logical operator for disjunction cannot be represented by the set-theoretic union, since the set-theoretical union of the two subspaces $P, Q$ will not be in general a subspace, thus, it is not an experimental proposition.
+Therefore, in QL one introduces the quantum logical disjunction as the the closed span of the subspaces $P, Q$, which is an experimental proposition.
+Such a statement corresponds to the ``smallest closed subspace containing both $P$ and $Q$'' (\cite{Bacciagaluppi:2009}, p.\ 55).
+\item The logical implication is defined by set theoretical inclusion: given two experimental propositions $p$ and $q$, 
+$p$ implies $q$, means that whenever one predicts $p$ with certainty, one can predict also $q$ with certainty, and this is equivalent to state that $p$ is a subset of $q$.
+This fact is particularly important since the authors showed that ``it is \emph{algebraically} reasonable to try to correlate physical qualities with subsets of phase-space'' and thus, ``physical qualities attributable to any physical system form a partially ordered system'' (\cite{vonNeumann:1936}, p.\ 828, notation adapted).
+\end{enumerate}
+
+In this manner we have defined a propositional calculus for quantum experimental propositions, which generates an orthocomplemented lattice.
+In what follows we analyze the motivations for such a quantum logic, clarifying why it is generally believed that classical propositional calculus cannot be applied in QM.
+
+\section{Quantum Disjunction and Distributivity Law}
+\label{Motivations}
+
+The first remarkable difference between QL and CL is the failure of the distributive law.
+It is a well-known fact that such a law holds in the propositional calculus underlying classical mechanics in virtue of the commutative algebra of classical observables.
+Referring to this, \cite{Giuntini:2002} claim that distributivity fails in quantum logic for two main reasons:
+\begin{itemize}
+\item the non-commutativity of quantum observables,
+\item the peculiar behavior of the quantum disjunction---which may be true although neither of the member is true in virtue of the linearity of the Schr\"odinger Equation (SE)---the other major deviation from classical logic.
+\end{itemize}
+\noindent Let us then have a closer look to these motivations for the failure of the distributivity law and more generally of CL in the quantum context.
+
+Contrary to the case of classical mechanics, QM relies on a non-commutative algebraic structure\footnote{For details on the mathematical structure of QM and its physical content the reader may refer e.g.\ to \cite{Sakurai1994}, and \cite{Griffiths:2014}.
+In this essay, it is assumed that the reader has some familiarity with the standard formalism of QM.}: quantum observables generally do not commute, meaning that for any pair of operators $A,B$ we have $[A,B]=AB-BA\neq0$.
+From a physical perspective this fact entails that if one performs a measurement of $A$ on a quantum system followed by a measurement of $B$, in general one will obtain different results inverting the order of these observations.
+Moreover, in virtue of the non-commutative structure of quantum theory, to measure an operator $C=A+B$ is not typically equivalent to measure $A$ and $B$ independently and then add the respective results, since $A,B$ will be in general incompatible (cf.\ \cite{David:2015}, p.\ 77).
+This algebraic fact of QM constitutes the formal basis to prove the Heisenberg uncertainty relation, a theorem of quantum theory which reflects the operational inability to simultaneously measure the values of incompatible operators with arbitrary precision.
+Such a result, in turn, is usually interpreted ontologically in the sense that quantum system do not instantiate definite properties in non-measurement situations (cf.\ \cite{Sakurai1994}).
+
+In addition, as we will explain in the remainder of this section, the failure of distributivity in quantum logic is due to another particular feature of this new propositional calculus, namely on the behavior of the quantum disjunction, which is due to the presence of quantum superpositions.
+Indeed, another remarkable difference between QM  and CM is that quantum systems can be in superposition states as a consequence of the linearity of the Schr\"odinger equation---the fundamental dynamical law of QM---which for a single particle reads:
+\begin{align}
+\label{SE}
+i\hbar\frac{\partial\psi}{\partial t}=\Big(-\frac{\hbar^2}{2m}\nabla_k^2+V\Big)\psi=H\psi,
+\end{align}
+\noindent where $H$ represents the Hamiltonian operator, defined as the sum of kinetic and potential energy of the system at hand.
+More specifically, this algebraic property of \eqref{SE} entails that if two wave functions\footnote{In QM the wave function of a system provides the maximal information available about it.} $\psi_1, \psi_2$ are both possible solutions of the same Sch\"odinger equation, then their linear combination (\emph{superposition})
+\begin{align}
+\label{SSE}
+\psi_s=\alpha\psi_{1}+\beta\psi_{2}
+\end{align} 
+is still a solution of the same SE---$|\alpha|^2,|\beta|^2$ (with $\alpha,\beta\in\mathbb{C}$) represent the probabilities to find the system in $\psi_1, \psi_2$ respectively.
+Notably, the new superposed state $\psi_s$ is also a consistent representation of the system.
+Referring to this,  \cite{deRonde:2016} explicitly say that in QL ``differently from the case in classical semantics, a quantum disjunction may be true even if neither of it members is true.
+This reflects, for example, the case in which we are dealing with a state such as that of a spin $1/2$ system which is in a linear combination of states up and down'', as in the example we are going to consider in order to explain the failure of distributivity in the quantum context.
+
+To claim that in quantum logic conjunction and disjunction are not distributive means that given three propositions $p, q$ and $r$ the logical law
+\begin{align}
+\label{DL}
+p\wedge(q\vee r)\longleftrightarrow(p\wedge q)\vee(p\wedge r)
+\end{align}
+
+\noindent does not hold.\footnote{In the context of quantum logic, distributivity must be replaced by a weaker law: $(p\wedge q)\vee(p\wedge r)\longrightarrow(p\wedge(q\vee r)).$}
+Let us explain this fact with a simple example taken from \cite{Giuntini:2002}.
+Suppose to consider a $1/2-$spin particle, let's say an electron and assume that this paritlce has spin up along the $x$ axis.
+From QM we know that there are only two possible spin states in which a particle can be found in each axis after a spin measurement, namely either in the up or in the down state.
+Furthermore, since spin operators along different axis do not commute, being incompatible observables---$[S_x, S_y]=[S_x, S_z]=[S_y, S_z]\neq0$---they cannot be simultaneously measured.
+Then, as a consequence of the Heisenberg uncertainty principle, ``both propositions ``$spin_y$ is up'' and ``$spin_y$ is down'' shall be strongly undetermined.
+However, the disjunction ``either $spin_y$ is up or $spin_y$ is down'' must be true'' (\cite{Giuntini:2002}, pp.\ 133-134), and the same holds for the case of spin in the $z$ direction.
+Therefore, if $p$ in \eqref{DL} represents the proposition ``the electron has $x$-spin up'', $q$ the proposition ``the electron has $y$-spin up'' and $r$  the proposition `the electron has $y$-spin down'', it is possible to see that the first part of \eqref{DL} express a true statement---i.e.\ $p\wedge(q\vee r)=1$---whereas $(p\wedge q)$ and $(p\wedge r)$ are both false in virtue of the incompatibility between the spin operators among different axes, implying that the second part of the distributive law asserts a false statement, i.e.\ $(p\wedge q)\vee(p\wedge r)=0$.
+Consequently, this law fails in the context of quantum theory, since we have $p\wedge(q\vee r)=1\longleftrightarrow(p\wedge q)\vee(p\wedge r)=0$.
+\vspace{2mm}
+
+In what follows we will stress that these motivations for QL are not sufficient to claim that CL is rejected by quantum theory.
 
@@
 
-The distributivity in quantum mechanics is ``weird'' because (1) it can be true even if none of the elements of the disjunction is true due to the linearity of the Schroendinger equation and (2) quantum observables do not commute.
+The distributivity in quantum mechanics is ``weird'' because (1) it can be true even if none of the elements of the disjunction is true due to the linearity of the Schr\"ondinger equation and (2) quantum observables do not commute.
 
@@
 
-In this case $p \wedge \hat{q} \nequiv \bot$ and $p \wedge \hat{r} \nequiv \bot$.
+In this case $p \wedge \hat{q} \not\equiv \bot$ and $p \wedge \hat{r} \not\equiv \bot$.
 Therefore the original claim was $p \wedge (\hat{q} \vee \hat{r})$ if and only if $(p \wedge  q) \vee (p \wedge r)$ which is not the classical logic distributive law.
@@
 \subsubsection{Classical example}
-\includegraphics[width=\columnwidth]{Balldrop.png}
+\begin{center}
+\includegraphics[scale=.3]{Balldrop.png}
+\end{center}
 
@@
 
-\section{Classical and quantum parallel}
-
-Classical state: $\rho(x,p)$
-
-Quantum state: $\psi(x)$
-
-Classical observable $f(x,p)$. Average value $<f> = \int f(x, p) \rho(x, p) dx dp$. Eigenstate $f(x, p) \rho(x, p) = f_0 \rho(x, p)$.
-
-Quantum observable $F$. Average value $<F> = \int \psi^\dagger(x) F \psi(x) dx$. Eigenstate $F \psi(x) = f_0 \psi(x)$.
-
-In both cases, average value equal to $f_0$ means that the contributions of all parts of the ensemble combine to the value $f_0$.
+%\section{Classical and quantum parallel}
+%
+%Classical state: $\rho(x,p)$
+%
+%Quantum state: $\psi(x)$
+%
+%Classical observable $f(x,p)$. Average value $<f> = \int f(x, p) \rho(x, p) dx dp$. Eigenstate $f(x, p) \rho(x, p) = f_0 \rho(x, p)$.
+%
+%Quantum observable $F$. Average value $<F> = \int \psi^\dagger(x) F \psi(x) dx$. Eigenstate $F \psi(x) = f_0 \psi(x)$.
+%
+%In both cases, average value equal to $f_0$ means that the contributions of all parts of the ensemble combine to the value $f_0$.
 Eigenstate $f_0$ means all parts of the ensemble contribute the same value $f_0$.
-
-For example, a classical eigenstate of $H = \frac{p^2}{2m} + \frac{1}{2} k x^2$ with eigenvalue $e_0$ is any function $\rho(x,p)$ that is different from zero only on the ellipses with energy $e_0$.
+%
+%For example, a classical eigenstate of $H = \frac{p^2}{2m} + \frac{1}{2} k x^2$ with eigenvalue $e_0$ is any function $\rho(x,p)$ that is different from zero only on the ellipses with energy $e_0$.
 That is, all elements of the distribution have the same energy.
-
-The statement ``the observable $f$ of the system is between $f_0$ and $f_1$'' is therefore unclear.
+%
+%The statement ``the observable $f$ of the system is between $f_0$ and $f_1$'' is therefore unclear.
 It could be taken to mean ``the average of observable $f$ of the system is between $f_0$ and $f_1$'' or ``the observable $f$ for each part of the ensemble is between $f_0$ and $f_1$''.
@@
 In the second case, we are identifying a set of possible functions based on their support.
-
-Now suppose that the $U$ and $V$ are two intervals.
+%
+%Now suppose that the $U$ and $V$ are two intervals.
 We have ``the average $<f>$ is in $U \cup V$'' is  equivalent to ``the average $<f>$ is in $U$'' $\vee$ ``the average $<f>$ is in $V$''.
@@
 But the lattice generated by such sets is still a $\sigma$-algebra, and therefore follows classical logic.
-
-To check/prove: neither algebras is a sub-algebra of the other; both sub-algebras of the norm-induced one.
+%
+%To check/prove: neither algebras is a sub-algebra of the other; both sub-algebras of the norm-induced one.
 
 \section{Math}
-
+\label{Math}
 In this section, we want to show how Hilbert spaces already are equipped with a mathematical structure that follows classical logic, namely the Borel algebra.
@@
 To show that some of these statements are indeed interesting, let us consider the following result:
+
 \begin{prop}
@@
 
-\section{Old notes}
-
-* We can always use the Borel algebra for our statements. This will include(?) all propositions of ``quantum logic''.
-
-It includes (?) also proposition on expectation values, say of position and momentum. Those will behave exactly like ``point particles''. It also includes (?) propositions on the support.
-
-Show that this is similar to classical distributions.
-
-Claim: in quantum mechanics the ``correct'' logical structure the is algebra of subspaces of the Hilbert space.
+\section{Discussion}
+\label{Discussion}
+%\section{Old notes}
+%
+%* We can always use the Borel algebra for our statements. This will include(?) all propositions of ``quantum logic''.
+%
+%It includes (?) also proposition on expectation values, say of position and momentum. Those will behave exactly like ``point particles''. It also includes (?) propositions on the support.
+%
+%Show that this is similar to classical distributions.
+%
+%
+%Claim: in quantum mechanics the ``correct'' logical structure the is algebra of subspaces of the Hilbert space.
 In this algebra (which is a $\cap$-structure) the meet functions differently than in Boolean logic.
 Therefore it does not follow classical logic.
-
-Resolution: that algebra does not include many statements that are interesting experimentally.
+%
+%Resolution: that algebra does not include many statements that are interesting experimentally.
 For example, it does not include statements about the expectation values which, in the end, are the result of experiments.
 (Cross-section?) One
-
-Technical view: the state space of quantum states is a separable Hilbert space; it is a Banach space (complete normed space) and admits a norm-induced topology; it is separable, so it admits a countable basis, so the topology is second countable; therefore the state space has a countably generated Borel-algebra.
+%
+%
+%Technical view: the state space of quantum states is a separable Hilbert space; it is a Banach space (complete normed space) and admits a norm-induced topology; it is separable, so it admits a countable basis, so the topology is second countable; therefore the state space has a countably generated Borel-algebra.
 This, like any set theoretic structure, follows Boolean algebra.
@@
 State-statements are all incompatible with each other: either we prepared one state or we prepared another.
-
-If a process with an initial and final state (the internal are disregarded), we have two lattices of propositions.
+%
+%If a process with an initial and final state (the internal are disregarded), we have two lattices of propositions.
 The sample space is the Cartesian product, the $\sigma$-algebra is the product.
@@
 These will naturally result in different lattices, but they will still be Boolean lattices.
-
-What changes are the probability spaces we can construct.
+%
+%What changes are the probability spaces we can construct.
 In the quantum case, the probability is not in general a probability measure and has to be consistent with the Born rule (i.e.
@@
 
-\bibliography{quantumlogic}
+\section{Conclusion}
+\label{conc}
+\vspace{5mm}
 
+\textbf{Acknowledgements:} withheld from draft.
+\clearpage
+
+\bibliographystyle{apalike}
+\bibliography{bibliography}
 \end{document}

### 2021-03-21 Articles/2020-QuantumLogic/Source/quantumlogic.tex
Another pass at the math section

@@
 \label{Math}
-In this section, we want to show how Hilbert spaces already are equipped with a mathematical structure that follows classical logic, namely the Borel algebra.
-We will see that the algebra typically associated with quantum logic is a subset of this structure.
-We will see that this classical structure contains other statements that are experimentally significant, like statements about the expectation values.
-Finally, we will show that the algebra of subspaces of classical distributions is also non-distributive in the same way that quantum logic is.
 
-Let us start with a brief summary of the basic mathematical structure of quantum logic.
-Each proposition for a Hilbert space $H$ correspond to a closed subspace.
-This forms a complete lattice under the following definitions: the ordering $p \leq q$ of the lattice (i.e.
-the implication $p \to q$) represents subspace inclusion, the conjunction $p \wedge q$ is the biggest subspaces contained by both, the disjunction $p \vee q$ is the smallest subspace that contains both, the negation $\neg p$ is the orthogonal subspace.
+In this section we present our main argument, which is that the state space of quantum mechanics and the space of classical probability distributions possess a parallel structure that allows one to apply commutative propositional calculus (i.e.
+``classical logic'') on the first and non-commutative propositional calculus (i.e.
+``quantum logic'') on the second.
+That is, both theories will have two lattices of propositions: a commutative one that follows classical logic and a non-commutative one that follows quantum logic.
 
-To understand how the previous structure fails to a be Boolean algebra, we note that by Stone's representation theorem any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.
-inclusion, intersection, union and complement).
-If we understand each subspace as a set of vectors, the disjunction $p \vee q$ is not the mere set union of the subspaces, but the span.
-This is why the above structure, even if it is a complemented $0-1$ lattice, it fails to be a Boolean alebgra.
+Note that the classical space we will be using is not phase-space itself, but the space of probability distributions over phase-space, which is a real Hilbert space and thus has a similar structure to the quantum counterpart.
+The central point is that many features of quantum logic stem not from quantum mechanics per-se, but from the fact that quantum mechanics is a probabilistic theory, which means a classical probabilistic theory will share the same features.
+Critically, we will see that the non-commutativity of the quantum logic lattice does not originate from non-commuting observables or superposition of quantum mechanics: it is a property of the lattice of subspaces of any vector space and, indeed, of the lattice of sugroups of any group more in general.
 
-However, a Hilbert space $H$ is a complete normed space.
-The norm induces a topology, which in turns induces a Borel-algebra.
-The Borel-algebra is a family of sets closed under countable intersection, union and negation.
-It therefore forms a Boolean algebra.
-Therefore propositions of the type ``the state is in set $A$'' for any Borel set $A$ follow the rules of classical logic.
+Mathematically, the classical commutative lattice will be provided by the Borel algebra, the smallest $\sigma$-algebra that contains the topology; the non-classical non-commutative lattice will be provided by the set of subgroups ordered by inclusion.
+We will see that the non-commutative lattice is a strict subset of the classical commutative one, meaning that everything that can be described in quantum logic can also be described in classical logic.
+Moreover, the commutative lattice will include propositions, like propositions on the expectation value of observables, that are of scientific interest.\footnote{In an experimental context, in fact, by ``measurement'' one typically refers to the statistical collection of many ``single takes''.
+Quantum logic describes the single-take measurements.
+For example, measurements of cross sections, the most typical type of measurement in the context of particle physics, are not part of the quantum lattice, and neither are measurements of the probabilities themselves.}
 
-Note that the lattice of quantum logic is formed by closed subspaces, which are closed sets and therefore Borel sets.
-Therefore all statements expressible in quantum logic (e.g.
-``we measured/prepared this quantity with the following value'') are also expressible with the Borel algebra.
-The converse, however, is not true: not all Borel sets correspond to propositions in quantum logic.
+While topologies and $\sigma$-algebras are foundational structures in mathematics for differential geometry,[CITE: Lee, Introduction to Smooth Manifolds] measure theory and probability theory,[CITE: Grimmet/Stirzaker, Probability and Random Processes] they do not seem well-known among physicists and philosophers of science, therefore we briefly summarize their definition, use and meaning.
+[CITE: Kelly, Logic of Reliable Inquiry - Carcassi/Aidala, Assumptions of Physics]
 
-To show that some of these statements are indeed interesting, let us consider the following result:
+We start with a set $X$ which can be thought to represent the set of all possible cases we need to distinguish: all quantum states in the quantum case; all possible probability distributions in the classical case; for a simpler example, the set of the real numbers for a continuous quantity.
+Both topologies and $\sigma$-algebras are collections of subsets of $X$.
+For example, the standard topology for the real numbers will contain all the open intervals $(a, b)$.
+The difference between a topology and a $\sigma$-algebra is what types of set operations one can take within them.
+
+A collection of subsets may seem quite an abstract concept, but it can be readily made concrete: any statement on the objects represented by $X$ can be represented by a set.
+For example, the open set $(a, b)$ will exactly correspond to the statement ``the value of the quantity is between $a$ and $b$''.
+It is therefore useful to think about both the topology and the $\sigma$-algebra as collections of statements about $X$ instead of sets.
+
+A $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, meaning that performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra.
+In terms of statements, this means that it is closed under countable conjunction, countable disjunction and complement: the $\sigma$-algebra can be thought as a Boolean algebra of statements.
+Note that this interpretation is exactly the one used in probability.
+In fact, a probability space is defined by three things: a sample space $\Omega$ which represents the set of all possible outcomes; a $\sigma$-algebra over $\Omega$ which represents all events, all propositions; and a measure that assigns a probability to each event.
+
+A topology, instead, is closed under arbitrary union and finite intersection only.
+In terms of the statements, it is closed under arbitrary disjunction and finite conjunction.
+A topology is therefore not a Boolean algebra, but a Heyting algebra.
+Heyting algebras play a central role in intuitionist logic (and in computer science) as they describe lattices of provable (or computable) statements in finite time.
+In physics, we can understand the topology to represent statements that are experimentally verifiable in finite time.[CITE again Kelly and Carcassi/Aidala] For example, the open interval $(a, b)$ is an open set precisely because the statement ``the value of the quantity is between $a$ and $b$'' is something we can verify experimentally.
+Note that the sets in a topology are called open sets precisely because they are a generalization of the open intervals.
+On the other hand, a set with only one real value $\{ x_0 \}$ is not an open set, it is not in the topology, because the statement ``the value of the quantity is exactly $x_0$'' is not verifiable experimentally due to infinite precision.
+
+Establishing a topology, then, means establishing what statements are assumed to be experimentally verifiable.
+Given a topology, its Borel algebra is the smallest $\sigma$-algebra that contains all sets in the topology.
+That is, it is the smallest Boolean algebra that contains all the verifiable statements associated to the topology.
+Topologies and $\sigma$-algebra are pervasive in physics, to the point that all physically interesting spaces are, directly or indirectly, topological spaces.
+And wherever we have a topology, we have a Borel algebra, which is a $\sigma$-algebra, a Boolean algebra.
+
+On complete normed spaces, such as Hilbert spaces, there is a standard way to construct the topology.
+Let us see first how it works in a standard three-dimensional Euclidean space.
+Given two points $x$ and $y$, we can write the square of distance $\epsilon^2 = |y - x|^2$ as the square of the norm of the vector between the two points.
+The statement ``the position is within $\epsilon$ meters of $x$'' corresponds to the set $U = \{y \; | \; |y - x|^2 < \epsilon^2 \}$ of all the points whose distance from $x$ is less than $\epsilon$.
+These are exactly the open sets that generate the topology.
+This works in the same way for all normed spaces.
+A Hilbert space, in particular, comes with an inner product which also defines a norm.
+We can define a distance $\epsilon^2 = |\psi - \phi|^2=\langle \psi - \phi , \psi - \phi \rangle$ between two states, and the statement ``the state of the system is within $\epsilon$ of $\psi$'' corresponds to the set $U = \{\phi \; | \; |\psi - \phi|^2 < \epsilon^2\}$.
+As we'll see later, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.
+We will have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
+In all normed spaces, the topology is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.\footnote{In the case of Hilbert spaces and quantum mechanics, the map between topology and experimental verification does not work perfectly because absolute modulus and phase are not physically meaningful.
+This can be fixed by using the corresponding projective space, which removes the overdetermination.
+We are not going to do this as projective spaces have their own peculiarities and are used less.}
+
+From this topology, one can construct the Borel algebra.
+The Borel algebra also contains the closed sets, which are defined to be the complement of open sets.
+Recall that, in quantum logic, the propositions corresponded to closed subspaces, which are closed sets and are therefore Borel sets.
+All the propositions of quantum logic are therefore readily found in the Borel algebra of the Hilbert space.
+The Borel algebra, which is a Boolean algebra that follows classical logic, contains also all the statements of quantum logic.
+This construction also applies identically for the classical cases, where we will also be able to talk about the set of all closed subspaces which will have a parallel structure.
+
+While all of this may seem daunting to someone who has never seen this material, we want to stress that these are very standard constructions.
+We are not taking some obscure framework from some fringe field of mathematics: these are the standard mathematical techniques used to study these spaces and can be found in any textbook.
+What we are doing here is bringing out their full physical significance, which is too often overlooked.
+Understanding the link between the mathematical structures and the physics they represent is crucial to clearing out possible misunderstandings.
+
+Now that we have briefly seen the definitions and basic constructions, we can gain more insights into these structures and what they represent.
+First, let us understand why the lattice of subspaces (i.e.
+quantum logic) fails to be a Boolean algebra.
+Stone's representation theorem tells us that any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.
+intersection, union and complement).
+For the lattice of subspaces, note that the disjunction $p \vee q$ does not correspond to the set union of the subspaces, but the span.
+In fact, the union of subspaces is not part of the lattice.
+This is why the lattice, even if it is a complemented $0-1$ lattice, fails to be a Boolean algebra: it does not use the standard set operations.
+
+Clearly, $\sigma$-algebras are Boolean algebras because they are exactly an algebra of sets with standard set operations.
+As we said before, the Borel algebra will contain the lattice of subspaces plus all the set operations between them, including the union.
+In a way, the lattice of subspaces is not Boolean because it does not contain enough sets, which are instead contained in the Borel algebra.
+
+One may ask: do these extra sets correspond to physically meaningful propositions? To answer that, let us first demonstrate the following proposition.
 
 \begin{prop}
-	Let $H$ be an Hilbert space and $\langle \, , \, \rangle$ be its inner product.
+	Let $H$ be a Hilbert space and $\langle \, , \, \rangle$ be its inner product.
 Let $X$ be a self-adjoint bounded linear operator.
-Let $F_X : H \to \mathbb{R}$ the map defined by $F_X(\psi) = \langle \psi , X \psi \rangle$.
+Let $F_X : H \to \mathbb{R}$ be the map defined by $F_X(\psi) = \langle \psi , X \psi \rangle$.
 Let $U \subseteq \mathbb{R}$ be a Borel set.
@@
 
-A proof would work along these lines.
-A linear operator between two normed spaces is continuous if and only if it is bounded, therefore our operator $X$ is continuous.
-Since the inner product is also continuous, $F_X$ is continuous.
-Since $F_X$ is continuous, the reverse image of an open set is an open set.
-Since a Borel set $U$ can be constructed from open sets using negation and countable union, the reverse image $F_X^{-1}(U)$ can also be constructed from open sets using negation and countable union, which means it is a Borel set.
+Proof.
+Let $X$ be a self-adjoint bounded linear operator.
+Recall that a linear operator is continuous if and only if it is bounded.
+Therefore $X$ is continuous.
+Note that $F_X$ is the composition of $X$ with the inner product.
+Recall that the inner product is a continuous function.
+Since $F_X$ is the composition of continuous functions it is also continuous.
+Recall that all continuous functions are also Borel measurable, therefore $F_X$ is Borel measurable.
+Let $U \subseteq \mathbb{R}$ be a Borel set.
+Since $F_X$ is Borel measurable, the reverse image $F_X^{-1}(U)$ is also a Borel set.
+This concludes the proof.
 
@@
 
-Borel sets allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
+Borel sets also allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
 This is NOT equivalent to the statement ``the particle is in the eigenstate $x_0$''.
@@
 
+We can go one step further.
 Note that all statements that characterize the position and momentum of the center of mass form a sub-algebra of the Borel algebra.
+Recall that the Ehrenfest theorem states that $m \frac{d}{dt}\langle x \rangle = \langle p \rangle$ and $ \frac{d}{dt}\langle p \rangle = - \langle V(x) \rangle$.
 If the potential for a quantum particle can be considered constant over the wave-function (i.e.
 $\langle V(x) \rangle = V(\langle x \rangle)$ ), the motion of the center of mass will reduce to the case of a point-particle.
-In fact, this is exactly how the point particle approximation arises in classical mechanics as well: when we study the motion of planets, cannonballs or specs of dust, we understand that these objects are not point-like, but they have an extent and a matter distribution.
-The point-particle approximation works insofar we are interested in the motion of the center of mass and the forces are such that the extent of the object does not matter.
-Naturally, if the orbit of a planet would be such that its distance is less than the radius of the sun, the point-particle approximation would fail spectacularly.
-In this regard, quantum mechanics works the same way.
+The Borel algebra contains as a sub-algebra the statements we need for (at least one form of) the classical limit.
+And this is a purely classical subalgebra, because the average position and average momentum can both be well defined.
+Moreover, note that this is not dissimilar to what happens in classical mechanics when, for example, we study planetary motion using only the position and momentum of the center of mass.
+The algebra of statements for the position and momentum of the center of mass can be independently described and studied.
+In this regard, the logical structure of the Borel algebra of quantum mechanics works in the same way.
 
 Now that we have shown that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us do the converse: show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
-As we said before, classical objects are not point-like.
-So let us consider the space of all possible matter distributions $\rho(x,p)$ over phase-space.
+Given that quantum mechanics is a probabilistic theory, let us compare it to a classical probabilistic theory.
+So let us consider the space of all possible probability distributions $\rho(x,p)$ over phase-space.
 We can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$.
 This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$.
-Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.\footnote{To be clear: we are not arguing that this construction is the most appropriate or useful.
+Note that $\langle \psi | F \psi \rangle = \int f(x,p) \rho(x,p) dx dp$ as one would expect.\footnote{To be clear, we are not arguing that this construction is the most appropriate or useful.
 Simply that it can be done.}  We can now take the algebra of subspaces and, as in the quantum case, it will be non-distributive: the disjunction of two subspaces will not be the set union, but the span.
 
-To simplify, let us assume that there are only two possible states $1$ and $2$.
-A state is identified by a two-valued distribution $[\rho_1, \rho_2]$, which we can take it to be either the probability of the particle being in each state or the amount of material in each state.
+To really understand the source of the non-commutativity, let us study the simplest possible example.
+Let us assume that there are only two possible classical states, $1$ and $2$.
+A state is identified by a two-valued distribution $[\rho_1, \rho_2]$, where $\rho_1$ is the probability to find the object in state 1 and $\rho_2$ is the probability to find the object in state 2.
 We note $[k,0]$ the subspace spanned by the first state only.
-This corresponds to the case where the whole system is in state $1$.
-Conversely, $[0,k]$ corresponds to the case where the whole system is in state $2$.
-The disjunction $[k,0] \vee [0,k]$ will correspond to the state spanned by both, which is the whole space $[k,j]$.
-We can also take the subspace $[k,k]$ which is the subspace spanned by vectors that have equal components in $1$ and $2$: the system is equally distributed in the two cases.
-Note that $[k,0]$, $[0,k]$ and $[k,k]$ are all pairwise disjoint: no pair have a vector in common apart from zero.
-Therefore their conjunctions $[k,0] \wedge [0,k] = [0,k] \wedge [k,k] = [k,k] \wedge [k,0] = [0,0]$ are the zero dimensional.
+This corresponds to the statement ``the system is certainly in state $1$'' since the probability for state $2$ is zero.
+Conversely, $[0,k]$ corresponds to the statement ``the system is certainly in state $2$''.
+The disjunction $[k,0] \vee [0,k]$ will correspond to the subspace spanned by both, which is the whole space $[k,j]$ where we can assign different non-zero probability to either case.
+We can also take the subspace $[k,k]$, the subspace spanned by vectors that have equal components in $1$ and $2$, which corresponds to the statement ``the system is equally likely to be in either state $1$ or $2$''.
+Note that $[k,0]$, $[0,k]$ and $[k,k]$ are all pairwise disjoint: no pairs have a vector in common apart from zero.
+Therefore their conjunctions $[k,0] \wedge [0,k] = [0,k] \wedge [k,k] = [k,k] \wedge [k,0] = [0,0]$ are zero dimensional.
 On the other hand, $[k,k]$ is a subspace of the whole space $[k,j]$ and therefore $[k,k] \wedge [k,j] = [k,k]$.
 Putting it all together, we have $[k,k] \wedge ( [k,0] \vee [0,k] ) = [k,k] \wedge [k,j] = [k,k]$.
-On the other hand, $( [k,k] \wedge [k,0] ) \vee ( [k,k] \wedge [0,k] ) = [0,0] \vee [0,0] = [0,0]$.
+But we also have $( [k,k] \wedge [k,0] ) \vee ( [k,k] \wedge [0,k] ) = [0,0] \vee [0,0] = [0,0]$.
 Therefore the lattice is not distributive.
 
-The lack of distributivity has nothing to do with the complexity of the Hilbert space or non-commuting observables.
-It's a property of the lattice of subspaces of a vector spaces.
-In fact, it's a property of the lattice on subgroups more in general.
-In the subgroup lattice, the disjunction is the span, so doing the disjuction first makes our set bigger during the calculation.
-Naturally, one may still argue that it is appropriate to focus on this lattice in quantum mechanics while it is not appropriate for classical mechanics.
-But one cannot argue that this difference is a fundamental difference in the mathematical or logical structure of the theories.
-Certainly, one cannot say that quantum mechanics requires a departure from classical mechanics when both theories follow similar structures.
+Note that during the discussion there was no mention as to whether $k$ and $j$ where real numbers, complex numbers or quaternions.
+The argument is formally the same in all cases.
+Therefore the lack of distributivity has nothing to do with the complexity of the Hilbert space or non-commuting observables.
+It's a property of the lattice of subspaces of any vector space and, more generally, of the lattice of subgroups of any group.
+In fact, consider the group of boost transformations in relativity.
+The boosts along one direction form a subgroup.
+If we take the subgroups of boost along $x$ and $y$ respectively, the smallest subgroup that contain both, the disjunction, is the subgroup of all boosts along any direction in the $x$-$y$ plane.
+As before, the disjunction is the span, so the resulting set is different from the union, and the union of all boosts along the x-axis and the y-axis is not a subgroup and therefore not contained in the lattice.
+Thus, this has the same non-commutative structure as quantum logic.
 
-The key difference between the algebras of statements, then, is not really about the type of the logical structure: it is its content in terms of statements.
+We want to be clear that we are not claiming that the lattice of subspaces of quantum logic is not of interest, or that it is of equal interest in both classical and quantum mechanics.
+We are pointing it out that the argument that quantum mechanics REQUIRES a radical departure from classical logic in untenable, given both theories follow similar structures.
+Such arguments have to be a lot more nuanced.
+Additionally, once we understand the similarity between the theories, we are in a much better position to understand the differences, which naturally are there.
+
+For example, quantum mechanics can only be expressed as a probabilistic theory and therefore ALWAYS requires a structure that in classical mechanics is not mandatory.
+In quantum mechanics, all subspaces of dimension one represent a pure state while in classical mechanics this is not the case.
 If we take two intervals $U_x$ and $U_p$ in position and momentum, and form the statement ``the state is wholly in $U_x$ and $U_p$'', meaning that the vector components are zero outside of those regions, in the classical case we always find at least one $\rho(x,p)$ that satisfies the bounds, while in quantum mechanics, if the bounds violate the uncertainty principle, no suitable $\psi(x)$ can be found.
-It is the content of the logic that changes.
+The point is that the differences are in the content of the logical structure, on the content of the statements, not on the type of structure.
 

### 2021-03-25 Overviews/Informational granularity/main.tex
Starting to finalize blueprints for informational granularity

--- 2021-02-22 Articles/2020-MeasureTheoryFoundation/main.tex
@@
-\documentclass{article}
+\documentclass[11pt]{article}
 \usepackage[utf8]{inputenc}
 
+\usepackage[margin=1in]{geometry}
 \usepackage{amsthm}
@@
 
-\title{Description granularity: a common foundation for geometry, measures and probability.}
-\author{gabriele.carcassi }
-\date{September 2020}
+\newtheorem{conj}[equation]{Conjecture}
+
+\title{Assumptions of Physics blueprint: information granularity as a common foundation for geometry, measures and probability.}
+\author{Gabriele Carcassi, Mark Greenfield, Christine Aidala }
+\date{\today}
 
@@
 \begin{abstract}
-    In this paper we present an order theoretic framework that we hope can be used as a common foundation for geometrical, measure theoretic and probabilistic concepts in the sciences.
-As in comparative probability, we start with a $\sigma$-complete Boolean algebra that represents all possible descriptions in a given context and impose a preorder.
+    This paper presents ideas and progress towards an order theoretic framework that can be used as a common foundation for geometrical, measure theoretic and probabilistic concepts in the sciences.
+This would address foundational questions such as: under what assumptions standard tools like calculus are physically meaningful?
+How do geometrical structures arise in physics?
+How can the idea of physical dimension be properly captured mathematically?
+The working theory is that geometrical and measure theoretic structures capture one and only one thing: how coarse or fine grained are the statements within a scientific theory.
+They should  be recovered from a single binary operator, that allows to compare two statements and say whether one provides a finer level of description than the other.
+Mathematically, fineness imposes a preorder on the algebra of statements of a scientific theory.
+Measure theoretical and geometrical notions will be recovered by picking a unit and quantifying the degree of precision in terms of that unit: ``the mass of the object is between 1 and 3 Kg'' is twice less precise than ``the mass of the object is between 0 and 1 Kg''.
+The objective is to find necessary and/or sufficient conditions for such constructions which can be given a straight forward physical motivation.
+%    As in comparative probability, we start with a $\sigma$-complete Boolean algebra that represents all possible descriptions in a given context and impose a preorder.
 In our case, the preorder does not represent likelyhood comparison, but rather whether one description is at a finer level of granularity than another.
@@
 
-	This work is part of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
+	This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
 \end{abstract}
@@
 
-This working draft summarizes the status of our work on a new approach to the foundations of geometry and measure theory and is intended to gather early feedback.
+\emph{This paper presents the goal and the current status of our research. It is intended to gather early feedback, including pointers to other relevant work.}
+
+The overall goal of Assumptions of physics is to to identify a handful of physical principles from which the basic laws can be rigorously derived.
+We have already established that topologies and $\sigma$-algebras can be recovered from requirements of experimental verifiability.\cite{AoPBook,AoPOverviewVerifiability} The next step is to understand what are the underpinning of geometrical and measure theoretic structures (including probability) which are layered on top of topologies and $\sigma$-algebras.
+As usual, a formal construction from scratch will force us to make explicit the physical requirements that are baked into the use of the common concepts and provide us with a framework that is, hopefully, still valid when those assumptions fail.
+
+Note that often in physics geometric structures are recovered from other geometric structures through the idea of emergence.
+For example, substantial effort is put in constructing theories where space-time emerge from other more fundamental structures.
+This does not solve the problem: it simply pushes it backwards.
+We are interested in the more fundamental problem of how the first geometric structure, and in fact any geometric structure, is defined.
+
+The idea is that while experimental verifiability keeps track of which statements in the theory are experimentally verifiable, informational granularity keeps track of which provides a finer level of description.
+\textbf{The working theory is that this, and only this, is what geometrical and measure theoretic structure capture: how coarse or fine grained are the statements within a scientific theory.} Finding the fundamental axioms associated to the notion of information granularity is therefore the central problem.
+
+%This working draft summarizes the status of our work on a new approach to the foundations of geometry and measure theory and is intended to gather early feedback.
 The goal is to rederive the usual concepts in geometry, measure theory and probability theory from a simpler more general structure, namely a preorder.
-This construction will force us to make explicit the physical requirements that are baked into the use of the common concepts and provide us with a framework that is, hopefully, still valid when those assumptions fail.
-
-The general idea is similar to what is done in comparative probability, yet we want the techniques extend from a single measure to a family of measures that allow us to define and compare sizes of sets at different scales (e.g.
+This construction will
+
+%The general idea is similar to what is done in comparative probability, yet we want the techniques extend from a single measure to a family of measures that allow us to define and compare sizes of sets at different scales (e.g.
 points, lines, surfaces, ...).
-We start with a $\sigma$-complete Boolean algebra $\tdomain$, a theoretical domain, that represents a set of statements, a set of descriptions, for a physical system in a given context.
-A Boolean algebra is also a lattice, and therefore already comes equipped with a partial order $\narrower$ such that $\stmt_1 \narrower \stmt_2$ if $\stmt_1 \OR \stmt_2 \equiv \stmt_2$.
+
+To give a sense of how this would work, we start with our notion of a theoretical domain $\tdomain$, which is a $\sigma$-complete Boolean algebra that represents a set of statements, a set of descriptions, for a physical system in a given context.
+Mathematically, a Boolean algebra is also a lattice, and therefore already comes equipped with a partial order $\narrower$ such that $\stmt_1 \narrower \stmt_2$ if $\stmt_1 \OR \stmt_2 \equiv \stmt_2$.
 This partial order indicates whether one statement gives a \textbf{narrower}, more specific, description than the other.
@@
 
-While we borrow many ideas from comparative probability, our goal cannot be achieved with a single measure.
+We found a few ideas from comparative probability useful, but our goal cannot be achieved with a single measure.
 A single measure can only compare objects with a finite measure.
@@
 Therefore we need to construct a family of measures, each identified by picking a particular statement that acts as a reference, as a unit.
-
-On the other hand, geometric structures do provide measure for different dimensionality by starting from a lowest definition (i.e.
+This provides a much richer structure, that is both interesting mathematically and more scientifically useful.
+
+As for geometry, geometric structures provide measures for different dimensionality by starting from a lowest definition (i.e.
 line distance for Riemannian and surface are for symplectic) and then constructing higher dimensional measures through products.
-While want to recover those cases as well, we want to be able to capture the physical idea of a unit system, which is not typically done.
-In that case, there will be dimensions that are not directly comparable (e.g.
+While want to recover those cases as well, this is not general enough.
+In physics there will be dimensions that are not directly comparable (e.g.
 pressure, volume, number of particles, chemical potential) but are still comparable under appropriate composition (i.e.
 pressure times volume with number of particles times chemical potential).
+We want to have a structure rich enough to capture these physical ideas (i.e.
+systems of units) which is not typically done.
 
@@
 
-In this overview we give a more detailed account of our goal and the status of our search.
-This will give a general idea of the mathematical issues, but it will also focus on the construction at a physical/conceptual level.
-It is important for our larger project that the mathematical detail proceeds in lock step with the conceptual and physical understanding.
-Later sections, instead, will concentrate on the bare formal aspects.
+We now give a more detailed overview of the goal and status.
+This will also cover the conceptual understanding and physical motivation, not just the mathematical aspects.
+Later sections will present both what has been established to work and what not to work in full mathematical detail.
+
+We were not able to find similar work in the literature.\footnote{If the reader has relevant suggestions, we would appreciate them.} The closest is in the context of comparative probability\cite{definetti, villegas} from which we have taken some mathematical ideas.
+The problem is that the preorder they typically use is essentially equivalent to the probability measure they recover, while the whole point in our case, as mentioned in the introduction, is that the preorder should give us a richer structure.
+
+As we will see later, the notion of infinitesimals plays an important role. Therefore there may be overlap with ideas in non-standard analysis, though we have not researched that literature.
 
@@
 Conceptually, this represents all the statements that can be associated with an experimental verification procedure.
-The full detail of why this is the correct abstraction is discussed in detail in our previous work.\footnote{See \url{https://assumptionsofphysics.org/book/}} It is a property of this type of spaces that each statement $\stmt \in \tdomain$ can be expressed as a disjunction of possible cases, of the  \textbf{possibilities} of the domain.
-Formally, each possibility is a minterm\footnote{A minterm is a disjunction in which each term appears once, either negated or not.} of the countable basis.
+The full detail of why this is the correct abstraction is discussed in detail in our previous work.\cite{AoPBook} It is a property of this type of spaces that each statement $\stmt \in \tdomain$ can be expressed as a disjunction of possible cases, of the  \textbf{possibilities} of the domain.\footnote{For example, ``the mass of the object is less than 1 Kg'' is the disjunction (i.e.
+the logical OR) of all the statements of the form ``the mass of the object is exact $x$ Kg'' where $0 \leq x < 1$.} Formally, each possibility is a minterm\footnote{A minterm is a disjunction in which each term appears once, either negated or not.} of the countable basis.
 The theoretical domain can then equivalently be represented by a $\sigma$-algebra.\footnote{This conceptually maps perfectly to what one has in probability: a sample space (i.e.
@@
 The first set of properties are the ones we must expect fineness to have in all cases.
-At this point, there are three, identified by DeFinetti for comparative probability, that we believe are necessary and that work as simple and intuitive starting points.
+At this point, there are three, identified by DeFinetti\cite{definetti} for comparative probability, that we believe are necessary and that work as simple and intuitive starting points.
 
@@
 This means that fineness must be uniquely bound by the impossibility and the certainty.
-Any contingent statement must be strictly between the two since there are some cases in which it will be true and others in which it will be false.
+Any contingent statement (i.e.
+one that can in principle be either true or false) must be strictly between the two since there are some cases in which it will be true and others in which it will be false.
 This is the first required property.\footnote{In terms of the $\sigma$-algebra, the full set is strictly the biggest set and the empty set is strictly the smallest set.}
@@
 
-While we are not convinced these are sufficient properties to define fineness, we are not aware of other properties that would make the cut.
+While we are not convinced these are sufficient properties to define fineness, we are also not found other properties that must be convincingly necessary. This is still an open area.
 
@@
 
-\textbf{Totality}.
+\textbf{Totality and physical dimensions}.
 If our objective is to construct a measure, totality is needed in the sense that every statement (i.e.
@@
 
+For example, in phase space (i.e.
+symplectic manifold) we would expect points to be comparable and equigranular $\eqgran$: each state gives the same description.
+This means that we can compare sets of finite points.
+Finite areas are also comparable to each other (through integration of the symplectic form), and are comparable to points.
+However, vertical lines (i.e.
+ranges in momentum alone) are not comparable to horizontal lines (i.e.
+ranges in position alone).
+Symplectic geometry, in fact, gives a size to areas and not to lines.
+
 Another counter example come from infinite ranges.
 Take the real line and consider the set of all positive numbers and the set of all negative numbers: should they be comparable?
-If we say they are equal, then that would tell us that there are more cases before 1 than after.
+If we say they are equal, then that would tell us that there are more cases before 0 than after.
 This means that the space is not symmetrical under translation.
@@
 However, once we pick the unit, we should be able to find the ``right'' set of statements on which to create the measure.
+Roughly speaking, two description of the same units are ``finitely comparable'' in the sense that one gives a finer description of the other of a finite factor.
+Descriptions of different unites are either ``infinitely comparable'' (i.e.
+areas are always bigger than lengths) or not comparable (i.e.
+position and momentum).
 One approach would be to ask the ``largest set of statements mutually comparable in fineness that contain the unit''.
@@
 \textbf{Boolean lattice}.
-Since equigranularity classes form a bounded partially ordered set, one can ask whether they should form a lattice.
+Granularity levels (i.e.
+equigranularity equivalence classes) form a bounded partially ordered set.
+One wonders whether they form a lattice.
 Physically this would make sense: we would be saying that given a set of statements $S \subseteq \tdomain$, we can always find a level of description that is the finest among all coarser descriptions than all the statements (i.e.
@@
 Conceptually, the addition of two level of precision $\dot{\stmt}_1 + \dot{\stmt}_2$ means returning the level of precision that can be achieved by taking the disjunction $\stmt_1 \OR \stmt_2$ of two incompatible statements $\stmt_1 \in \dot{\stmt}_1$ and $\stmt_2 \in \dot{\stmt}_2$.
-This, of course, is not possible if the statements are ``bigger'' than ``half'' the space.
-A more order theoretical definition would be to take all possible disjuctions of $\stmt_1 \OR \stmt_2$ such that $\stmt_1 \in \dot{\stmt}_1$ and $\stmt_2 \in \dot{\stmt}_2$ and return the coarsest level of description.
+This, of course, is not possible if the statements are ``bigger'' than ``half'' the space (e.g.
+$\stmt_1 \eqgran \stmt_2$, $\stmt_1 \comp \stmt_2$ and $\stmt_1 \OR \stmt_2 \equiv \certainty$).
+A more order theoretical definition would be to take all possible disjuctions of $\stmt[v]_1 \OR \stmt[v]_2$ such that $\stmt[v]_1 \in \dot{\stmt}_1$ and $\stmt[v]_2 \in \dot{\stmt}_2$ and return the coarsest level of description.
 That is, the sum is the least upper bound in fineness of all possible disjunctions.
@@
 It is not clear whether the axioms above are enough to show that the addition always exists.
-It may be the existence of the addition, or a premise that leads to it, needs to be added.
+It may be that the existence of the addition, or a premise that leads to it, needs to be added.
 It is not clear whether, physically, such addition always makes sense, particularly in the context of incomparable descriptions (e.g.
@@
 In our framework, we have already a notion of independence that is defined on statements and logic operation, which is based on the idea of whether fixing the truth value of some statements will also fix the truth value of other statements.
+
+\textbf{Missing axioms}.
+In many of our proofs, we ran into an issue of not being able to guarantee existence of certain statements.
+Several of the definitions we wanted included statements like ``say x holds if there exist statement y such that w." However, we have no axioms that guarantee any existence of any statements beyond $\certainty$ and $\impossibility$.
+In the case of an infinite collection of possibilities, we have no way to guarantee existence of statements of various levels of fineness when all statements in question are compatible with infinitely many possibilities.
+
+To summarize, we may want to add something about existence of statements in equigranularity classes so as to have no ``gaps" (e.g.
+the order on $\tdomain/_{\eqgran}$ is complete).
+We may also want to say something about existence of enough statements representing each equivalence class so that sum and product can be well-defined.
+For example, if $\stmt_1\finer\stmt_2$, we want to say that there exists $\stmt_1'$ with $\stmt_1'\eqgran\stmt_1$ and $\stmt_1'\narrower\stmt_2$, as well as $\stmt_1''$ such that $\stmt_2\AND\stmt_1''$ is minimal in some sense ($\stmt_2\ncomp\stmt_1''$ if they do not fill the space).
+This is not at all guaranteed by the decomposition of statements into possibilities, since cardinality does not help us here.
+
+\textbf{Bigger and smaller halfs}.
+One thing that appears to be missing is the following.
+Consider the following domain with four possibilities $\{x_1, x_2, x_3, x_4\}$.
+Each statement is the disjunction of a set of possibilities and let us use the lower indexes to indicate them.
+For example, $\stmt_{12} = x_1 \OR x_2$.
+Now, suppose $\stmt_{12} \broader \stmt_{34}$ and $\stmt_{14} \broader \stmt_{23}$.
+The two pairs are incompatible (disjoint) statements whose conjunction is the certainty.
+In a sense $\stmt_{14}$ is more than half the certainty while $\stmt_{34}$ is less than half the certainty.
+So, we would expect $\stmt_{14} \broader \stmt_{34}$.
+Yet, our axioms do not guarantee that.
+To prove that, we wrote a program that constructed all possible relationships given the axioms, and that relationship was not derived.
+
+One may think we are then able to create a nonsensical situation, where $\stmt_{34} \broader \stmt_{14}$.
+What happens is that now we can apply offset monotonicity multiple times like so: $\stmt_{34} \broader \stmt_{14} \implies \stmt_{3} \broader \stmt_{1} \implies \stmt_{23} \broader \stmt_{12}$.
+Therefore $\stmt_{12} \broader \stmt_{34} \broader \stmt_{14} \broader \stmt_{23} \broader \stmt_{12}$.
+Which means $\stmt_{12} \eqgran \stmt_{34} \eqgran \stmt_{14} \eqgran \stmt_{23}$.
+Therefore, while we can't derive the fact that the bigger halfs are coarser than smaller halfs, if we try to force a smaller half to be coarser than a bigger half we actually get equal halfs.
+That is: if the two halfs are comparable, then we must have at least that the bigger half is coarser than the smaller half.
+This is interesting because, in this case, the mere assumption of comparability force at least a specific relationship.
+
+So the real issue is: does it make sense for those halfs to not be comparable? It is still not clear one way or the other.
 
@@
 This is tightly linked with a notion of infinitesimal (i.e.
-the ordering over the level of granularity is non-archemedian, while the real numbers are).
-The idea is to proceed in the following way:
+the ordering over the level of granularity is non-archemedian, while the real numbers are) which we will see in the next subsection.
+The address this collapse in a controlled way, the idea is to proceed in the following way:
 \begin{enumerate}
@@
 
-The Archemedian property is typically discussed in terms of linearly ordered groups.
-\footnote{See \url{https://en.wikipedia.org/wiki/Archimedean_property} for an introduction.} Roughly speaking, if we pick an element $y$ an element $x$ is infinitesimal with respect to $y$ if $\sum\limits_{i=1}^n x < y$ for any $n$.
-The group is Archemedian if no element is infinitesimal with respect to the other.
-We need to extend this notion to our ordering in fineness.
-
-The first issue, then, is to find the correct definition to extend this concept on ordering.
-The addition discussed about (i.e.
-roughly the disjoint union of statements) should play the role of the group operation.
+\textbf{Overview}.
+The measure zero sets are related to the notion of infinitesimals and the Archimedean property.
+The Archimedean property is typically discussed in terms of linearly ordered groups and fields.
+Roughly speaking, if we pick an element $y$ an element $x$ is infinitesimal with respect to $y$ if $\sum\limits_{i=1}^n x < y$ for any $n$.
+The group is Archimedean if no element is infinitesimal with respect to another.
+
+The (positive) real numbers are Archimedean.
+If we take a lattice of sets and use disjoint union as addition, the lattice is not Archimedean.
+A measure will therefore need a non-Archimedean structure (i.e the sets) into an Archimedean one (i.e.
+the positive reals).
+Infinitesimals therefore need to be mapped to zero.
+Note that for the measure to be countably additive, we need to have that the countable sum of infinitesimals is still infinitesimals.
+This will not be true in general: a set with finitely many integers is infinitesimal with respect to the set of all integers; yet the set of all integers is the disjoint union of countably many finite sets of integers.
+
+Note that on Page 21 de Finetti\cite{definetti} specifically says that $P(E) \geq P(E’)$ implies $E \geq E’$ but not vice-versa.
+And that we can have $P(A)=P(B)=0$ even if $A<B$ if $A$ is impossible and $B$ is not.
+So, the ordering he has in mind is fineness for us.
+And then he says that this means that the 4 postulates give a non-Archimedean structure.
+But you can make it Archimedean by ignoring the probabilities that are “infinitesimally small”: those that multiplied by a number, however big, will never become the certainty.
+On the other hand, Villegas\cite{villegas} starts already with an ordering where the infinitesimals are equal to the empty set.
+This maps to $\lequ$ defined later, where the comparison is done up to infinitesimals.
+
+\textbf{Definition of infinitesimals}.
+The first issue is to find the correct definition to extend this concept on ordering.
+The addition discussed above (i.e.
+the disjoint union of statements) should play the role of the group operation.
 Conceptually, a statement $\stmt$ is infinitesimal with respect to another $\stmt[u]$ if for any $n$ we can find a finite set of equigranular incompatible statements whose disjunction is finer than $\stmt[u]$.
@@
 The main issue here is that, since the ordering is potentially not total, it is not clear how one can prove the existence of said sets of statements even in the simplest cases.
-
+Therefore we have to find a definition that works well also in the case of non-comparability.
+
+\textbf{Definition of infinitesimals}.
 Assuming those issues are solved, infinitesimals come in two types: countable and uncountable.
-A countable infinitesimal will be such that the disjoint conjunction of countably many equigranular statements will be bigger than the unit.
-This is the case where, in the end, the measure will only be finitely additive: infinitely many zero measure elements will some to a finite element.
+A countable infinitesimal will be such that the disjoint conjunction of countably many equigranular statements may be bigger than the unit.
+This is the case where, in the end, the measure will only be finitely additive: infinitely many zero measure elements will sum to a finite element.
 If all infinitesimal with respect to the unit are uncountable, instead, countably many zero measure elements will still form a zero measure element.
@@
 All sets of measure zero are uncountable infinitesimals with respect to the unit: since the measure is countably additive, any disjoint union of countably many measure zero sets is going to be measure zero, therefore it will not be as big as the unit.
-Natuarally, if we take an uncountably many sets (i.e.
-all the singletons) we would cover the whole unit.
+Naturally, if we take a uncountably many sets (i.e.
+all the singletons) we can cover the whole unit.
 Conversely, the full real line can be covered by countably many sets of unit size.
 While the unit interval is infinitesimally small with the respect with the whole line, it is a countable infinitesimal.
-This definition of ``finiteness'' would map to physical requirements much better than, for example, the topological notion of compactness.
+This definition of ``finiteness'' may map to physical requirements much better than, for example, the topological notion of compactness.
 
@@
 
-One issue will be the problem of countable infinitesimals, which are incompatiable with a countable measure.
+\textbf{Countable addittivity}.
+One issue will be the problem of countable infinitesimals, which are incompatible with a countable measure.
 Therefore one of the assumptions in comparative probability will effectively exclude the case.
 We need to understand how that works, so that we can adapt the proofs and allow for only finitely additive measures.
-
-The other issue is that the measure will not be, in general, on the whole domain $\tdomain$, but on a subset $\tdomain_{\stmt[u]}$ that must be totally comparable with $\lequ$.
+The hope is that if we pick a unit for which all infinitesimals are uncountable, then one can demonstrate the requirements for the previous theorems
+
+\textbf{Measure domain}.
+The other issue is that the measure will not be, in general, on the whole domain $\tdomain$, but on a subset $\tdomain_{\stmt[u]}$ that must be totally comparable under $\lequ$.
 It is not clear how to recover this set.
@@
     \item Prove that if $\stmt[v] \narrower \stmt[u]$ then $\mu_{\stmt[u]}(\stmt[w]) = \mu_{\stmt[u]}(\stmt[v]) \mu_{\stmt[v]}(\stmt[w]) $ for all $\stmt[w] \narrower \stmt[v]$.
-To achive this, we will need to prove that the infinitesimals with respect to $\stmt[u]$ are exactly the infinitesimals with resepect to $\stmt[v]$.
+There are going to be two cases.
+If $\stmt[v]$ is infinitesimal with respect to $\stmt[u]$ then $\mu_{\stmt[u]}(\stmt[w])$ is zero and $\mu_{\stmt[u]}(\stmt[v])$ is zero.
+If not, we know that the infinitesimals with respect to $\stmt[u]$ are exactly the infinitesimals with resepect to $\stmt[v]$, so the equivalence classes of $\equ$ and $\equ[v]$ are the same.
+At that point, one should be able to use the properties of the measures to prove the rest.
     \item Given two statements $\stmt[u] \narrower \stmt[v]$, we extend the measure on $\stmt[u]$ by defining $\mu_{\stmt[u]}(\stmt[v]) = 1/ \mu_{\stmt[v]}(\stmt[u])$.
@@
 
+\subsection{Basic fineness definitions}
+
 \begin{defn}
@@
 
-The fact that the $\sigma$-algebra is generated by a countable set means it is separable. It also means that each element can be constructed as a disjunction of a set of possibilities.
+The fact that the $\sigma$-algebra is generated by a countable set means it is separable.
+It also means that each element can be constructed as a disjunction of a set of possibilities.
+For the full definitions, see \cite{AoPBook}.
 
@@
 
+\begin{defn}
+    We say $\stmt_1$ is \textbf{equigranular} to $\stmt_2$ (noted $\stmt_1 \eqgran \stmt_2$) if $\stmt_1 \finer \stmt_2$ and $\stmt_2 \finer \stmt_1$.
+We say $\stmt_1$ is \textbf{strictly finer} than $\stmt_2$ (noted $\stmt_1 \sfiner \stmt_2$) if $\stmt_1 \finer \stmt_2$ but not $\stmt_2 \finer \stmt_1$.
+\end{defn}
+
 \begin{prop}\label{fineness_properties}
@@
     \begin{enumerate}
-        \item if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \sfiner \stmt_2$ if and only if $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$ (these two tell us that monotony can be extend to strict fineness and equigranularity)
+        \item if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \sfiner \stmt_2$ if and only if $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$ (these two tell us that offset monotonicity can be extend to strict fineness and equigranularity)
         \item if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \eqgran \stmt_2$ if and only if $\stmt \OR \stmt_1 \eqgran \stmt \OR \stmt_2$
@@
 \end{prop}
-\begin{remark}
-Everything is proved without using totality, so that (if we remove it) everything follows.
-\end{remark}
 
@@
 Suppose $\stmt_1 \sfiner \stmt_2$, then $\stmt_1 \finer \stmt_2$ and $\stmt_1 \ncoarser \stmt_2$.
-By monotony we have $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \ncoarser \stmt \OR \stmt_2$.
+By offset monotonicity we have $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \ncoarser \stmt \OR \stmt_2$.
 Therefore $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$.
 Conversely, if $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$ then $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \ncoarser \stmt \OR \stmt_2$.
-By monotony $\stmt_1 \finer \stmt_2$ and $\stmt_1 \ncoarser \stmt_2$, which means $\stmt_1 \sfiner \stmt_2$.
+By offset monotonicity $\stmt_1 \finer \stmt_2$ and $\stmt_1 \ncoarser \stmt_2$, which means $\stmt_1 \sfiner \stmt_2$.
 
 For 2, now suppose $\stmt_1 \eqgran \stmt_2$, then $\stmt_1 \finer \stmt_2$ and $\stmt_1 \coarser \stmt_2$.
-By monotony we have $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \coarser \stmt \OR \stmt_2$.
+By offset monotonicity we have $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \coarser \stmt \OR \stmt_2$.
 Therefore $\stmt \OR \stmt_1 \eqgran \stmt \OR \stmt_2$.
 Conversely, if $\stmt \OR \stmt_1 \eqgran \stmt \OR \stmt_2$ then $\stmt \OR \stmt_1 \finer \stmt \OR \stmt_2$ and $\stmt \OR \stmt_1 \coarser \stmt \OR \stmt_2$.
-By monotony $\stmt_1 \finer \stmt_2$ and $\stmt_1 \coarser \stmt_2$, which means $\stmt_1 \eqgran \stmt_2$.
+By offset monotonicity $\stmt_1 \finer \stmt_2$ and $\stmt_1 \coarser \stmt_2$, which means $\stmt_1 \eqgran \stmt_2$.
 
@@
 We have $\certainty \equiv \NOT \stmt_2 \OR \stmt_2 \coarser \NOT \stmt_2 \OR \stmt_1$.
-By monotony we have $\stmt_2 \coarser \stmt_1$.
+By offset monotonicity we have $\stmt_2 \coarser \stmt_1$.
 
@@
 For 6, note that $\stmt_1 \equiv (\stmt_1 \AND \NOT \stmt_2) \OR (\stmt_1 \AND \stmt_2)$, $\stmt_2 \equiv (\NOT \stmt_1 \AND \stmt_2) \OR (\stmt_1 \AND \stmt_2)$ with $\stmt_1 \AND \stmt_2 \ncomp \stmt_1 \AND \NOT \stmt_2$ and $\stmt_1 \AND \stmt_2 \ncomp \NOT \stmt_1 \AND \stmt_2$.
-By monotony, $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \AND \NOT \stmt_2 \finer \NOT \stmt_1 \AND \stmt_2$.
+By offset monotonicity, $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \AND \NOT \stmt_2 \finer \NOT \stmt_1 \AND \stmt_2$.
 
@@
 For 12, suppose $\stmt \finer \impossibility$. By unique boundedness, $\stmt$ can neither be contingent nor certain. Therefore $\stmt$ must be impossible.
-
 \end{proof}
@@
 
-\begin{defn}
-    We say $\stmt_1$ is \textbf{equigranular} to $\stmt_2$ (noted $\stmt_1 \eqgran \stmt_2$) if $\stmt_1 \finer \stmt_2$ and $\stmt_2 \finer \stmt_1$.
-\end{defn}
-
 \begin{prop}
-Other properties useful properties:
+Other properties satisfied by fineness:
 \begin{enumerate}
+    \item if $\stmt[s]_1 \ncomp \stmt[s]_2$ and $\stmt[v]_1 \ncomp \stmt[v]_2$, then $\stmt_1 \finer \stmt[v]_1$ and $\stmt_2 \finer \stmt[v]_2$ imply  $\stmt_1 \OR \stmt_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
     \item (monotony as in Villegas) if $\stmt[v]_1 \ncomp \stmt[v]_2$, then $\stmt_1 \finer \stmt[v]_1$ and $\stmt_2 \finer \stmt[v]_2$ imply  $\stmt_1 \OR \stmt_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
-    \item if $\stmt_1 \ncomp \stmt[v]_1$, $\stmt_2 \ncomp \stmt[v]_2$ and $\stmt[v]_1 \eqgran \stmt[v]_2$, then $\stmt_1 \finer \stmt_2$ if and only if $\stmt_1 \OR \stmt[v]_1 \finer \stmt_2 \OR \stmt[v]_2$.
+    \item if $\stmt_1 \ncomp \stmt_2$, $\stmt[v]_1 \ncomp \stmt[v]_2$ and $\stmt_2 \eqgran \stmt[v]_2$, then $\stmt_1 \finer \stmt[v]_1$ if and only if $\stmt_1 \OR \stmt_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
 \end{enumerate}
@@
 
-\begin{proof}
-
-\end{proof}
+\begin{center}
+\includegraphics[scale=.2]{VillegasDiagram.jpg}
+\end{center}
+
+\begin{proof}
+For 1, using the diagram as an aid, let us decompose the four statements into the disjunction of eight pairwise incompatible statements. We have
+\begin{itemize}
+    \item $\stmt[v]_1 \equiv \stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_8 $
+    \item $\stmt[v]_2 \equiv \stmt[w]_4 \OR \stmt[w]_5 \OR \stmt[w]_6 $
+    \item $\stmt[s]_2 \equiv \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_4 $
+    \item $\stmt[s]_1 \equiv \stmt[w]_6 \OR \stmt[w]_7 \OR \stmt[w]_8 $
+\end{itemize}
+where
+\begin{itemize}
+    \item $\stmt[w]_1 \equiv \stmt[v]_1 \AND \NOT \stmt[s]_1 \AND \NOT \stmt[s]_2 $
+    \item $\stmt[w]_2 \equiv \stmt[v]_1 \AND \stmt[s]_2 $
+    \item $\stmt[w]_3 \equiv \stmt[s]_2 \AND \NOT \stmt[v]_1 \AND \NOT \stmt[v]_2 $
+    \item $\stmt[w]_4 \equiv \stmt[v]_2 \AND \stmt[s]_2 $
+    \item $\stmt[w]_5 \equiv \stmt[v]_2 \AND \NOT \stmt[s]_1 \AND \NOT \stmt[s]_2 $
+    \item $\stmt[w]_6 \equiv \stmt[v]_2 \AND \stmt[s]_1 $
+    \item $\stmt[w]_7 \equiv \stmt[s]_1 \AND \NOT \stmt[v]_1 \AND \NOT \stmt[v]_2 $
+    \item $\stmt[w]_8 \equiv \stmt[v]_1 \AND \stmt[s]_1 $
+\end{itemize}
+
+Since all $\{ \stmt[w]_i \}$ are pair-wise incompatible, offset monotonicity allows to remove or add the same statement on both sides of a fineness relationship. Therefore
+\begin{itemize}
+    \item $\stmt[v]_1 \coarser \stmt[s]_1 \implies  \stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_8 \coarser \stmt[w]_6 \OR \stmt[w]_7 \OR \stmt[w]_8 \implies \stmt[w]_1 \OR \stmt[w]_2 \coarser \stmt[w]_6 \OR \stmt[w]_7 \implies \stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_5 \coarser \stmt[w]_5 \OR \stmt[w]_6 \OR \stmt[w]_7$
+    \item $\stmt[v]_2 \coarser \stmt[s]_2 \implies  \stmt[w]_4 \OR \stmt[w]_5 \OR \stmt[w]_6 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_4 \implies \stmt[w]_5 \OR \stmt[w]_6 \coarser \stmt[w]_2 \OR \stmt[w]_3 \implies \stmt[w]_5 \OR \stmt[w]_6 \OR \stmt[w]_7 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_7$
+    \item $\stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_5 \coarser \stmt[w]_5 \OR \stmt[w]_6 \OR \stmt[w]_7 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_7 \implies \stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_5 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_7 \implies \stmt[w]_1 \OR \stmt[w]_5 \coarser \OR \stmt[w]_3 \OR \stmt[w]_7$
+\end{itemize}
+Finally $\stmt[w]_1 \OR \stmt[w]_5 \coarser \stmt[w]_3 \OR \stmt[w]_7 \implies \stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_4 \OR \stmt[w]_5 \OR \stmt[w]_6 \OR \stmt[w]_8 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_4 \OR \stmt[w]_6 \OR \stmt[w]_7 \OR \stmt[w]_8 \implies \stmt[v]_1 \OR \stmt[v]_2 \coarser \stmt[s]_1 \OR \stmt[s]_2$.
+
+For 2, let $\hat{\stmt[s]}_2 = \stmt[s]_2 \AND \NOT \stmt[s]_1$.
+We have $\stmt[s]_1 \ncomp \hat{\stmt[s]}_2$ and $\hat{\stmt[s]}_2 \finer \stmt[s]_2 \finer \stmt[v]_2$ since $\hat{\stmt[s]}_2 \narrower \stmt[s]_2$.
+By 1, we have $\stmt[s]_1 \OR \hat{\stmt[s]}_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
+But since $\stmt[s]_1 \OR \hat{\stmt[s]}_2 \equiv \stmt[s]_1 \OR (\stmt[s]_2 \AND \NOT \stmt[s]_1) \equiv (\stmt[s]_1 \OR \stmt[s]_2) \AND (\stmt[s]_1 \OR \NOT \stmt[s]_1) \equiv \stmt[s]_1 \OR \stmt[s]_2$, we have $\stmt[s]_1 \OR \stmt[s]_2 \finer \stmt[v]_1 \OR \stmt[v]_2$..
+
+\end{proof}
+
+\begin{conj}
+Other properties that show how the most you can broaden something within an equigranular equivalence class is by a disjunction with something incompatible.
+
+\begin{itemize}
+	\item if $\stmt_1 \ncomp \stmt_2$, $\stmt[v]_1 \ncomp \stmt[v]_2$ and $\stmt_2 \eqgran \stmt[v]_2$, then $\stmt_1 \finer \stmt[v]_1$ if and only if $\stmt_1 \OR \stmt_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
+
+	\item if $\stmt[v]_1\ncomp\stmt[v]_2$, $\stmt[v]_2\eqgran\stmt_2$, and $\stmt_1\finer\stmt[v]_1$. Then $\stmt_1\OR\stmt_2\finer\stmt[v]_1\OR\stmt[v]_2$.
+\end{itemize}
+
+\end{conj}
 
@@
 \begin{defn}
-    Let $\tdomain_{/\eqgran}$ be the granularity equivalence classes. Let $(\cdot)_{/\eqgran} : \tdomain \to \tdomain_{/\eqgran}$.
+    Let $\tdomain_{/\eqgran}$ be the equivalence classes of equigranularity.
+We call \textbf{granularity level} an equivalence class of equigranularity.
+Let $(\cdot)_{/\eqgran} : \tdomain \to \tdomain_{/\eqgran}$ the map from a statement to its granularity level, which we also note $\dot{\stmt} = {\stmt}_{/\eqgran}$.
+We say $\dot{\stmt}$ is the granularity level of $\stmt$.
 \end{defn}
@@
 It will be crucial that two equigranular statements can be incompatible, and this in contradiction with $(\cdot)_{/\eqgran}$ being a lattice homomorphism.
-In fact, suppose $\tdomain_{/\eqgran}$ is a lattice and let $\wedgedot$ be its meet.
-Suppose that $\stmt_1 \eqgran \stmt_2 \nequiv \impossibility$ and $\stmt_1 \ncomp \stmt_2$.
-Then ${\stmt_1}_{/\eqgran} = {\stmt_2}_{/\eqgran} \neq \impossibility_{/\eqgran}$.
-Now suppose that $(\cdot)_{/\eqgran}$ is a homomorphism.
-Then $\impossibility_{/\eqgran} = {(\stmt_1 \AND \stmt_2)}_{/\eqgran} = {\stmt_1}_{/\eqgran} \wedgedot {\stmt_2}_{/\eqgran} = {\stmt_1}_{/\eqgran} \wedgedot {\stmt_1}_{/\eqgran} = {\stmt_1}_{/\eqgran}$, which contradicts the premise.
+Let us note $\dot{\stmt} = {\stmt}_{/\eqgran}$ the level of granularity of $\stmt$.
+Now suppose that $\stmt_1 \eqgran \stmt_2 \nequiv \impossibility$ and $\stmt_1 \ncomp \stmt_2$.
+Then $\dot{\stmt}_1 = \dot{\stmt}_2 \neq \dot{\impossibility}$.
+Now suppose that $\stmt \to \dot{\stmt}$ is a lattice isomorphism.
+Then $\dot{\impossibility} = (\stmt_1 \AND \stmt_2)_{/\eqgran} = \dot{\stmt}_1 \wedgedot \dot{\stmt}_2 = \dot{\stmt}_1 \wedgedot \dot{\stmt}_1 = \dot{\stmt}_1$, which contradicts the premise.
 \end{remark}
@@
 \end{defn}
+
+It would be nice to be able to prove the following either true or false.
+
+\begin{conj}\label{uniformconsequence}
+    Let $\tdomain$ be uniform.
+Let $\stmt, \stmt_1 \in \tdomain$ such that $\stmt \sfiner \stmt_1$.
+Then there exist $\stmt_2 \in \tdomain$ such that $\stmt_2 \narrower \stmt_1$ and $\stmt \eqgran \stmt_2$.
+\end{conj}
+
+[TODO: clean-up what follows]
+
+The start of a proof....
+Let $S = \{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$. We will find a maximal element with respect to fineness (OR narrowness??).
+
+The difficulty here is that for Zorn we need to find an upper bound of an arbitrary chain which also lies in $S$.
+But such an upper bound is entirely unclear on how to obtain without using the proposition to begin with.
+Uniformity is difficult to make much use of for infinite domains (beyond that $S$ is nonempty).
+
+This set $S$ is nonempty, since any possibility narrower than $\stmt_1$ must also be finer than $\stmt$ by uniformity, and the statement $\stmt_1$ must not be the impossibility since it is strictly coarser than the statement $\stmt$.
+
+Now, given any chain (with respect to narrowness), consider its disjunction. But again this may be coarser than $\stmt$!
+
+Note: problem with chains for fineness.
+Taking the disjunction of a chain with respect to fineness can yield something much coarser (i.e.
+outside of $\{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$) if they do not overlap.
+
+Idea: given chains in fineness, can we construct new chains which are, at each level, equigranular, but now they are also chains in narrowness?
+
+Lemma: If $C$ is a chain in fineness and $\stmt$ is such that $\stmt[t]\finer\stmt$ for all $\stmt[t]\in C$, then there exists a set $N$ and a bijection $f:C\to N$ such that $N$ is a chain in narrowness, $\OR N \narrower \stmt$, and for all $\stmt[t]\in C$, $\stmt[t]\eqgran f(\stmt[t])$.
+Use this to move chains in fineness over to chains in narrowness and then can find maximal elements in the set $S$ above ``up to equigranular equivalence classes".
+
+Problem: this is actually much harder. Prop 6 is the case of this Lemma for a chain of length 2 and the top of the chain being equal to $\stmt$.
+
+\subsection{Infinitesimals}
+
+The following is an attempt to define infinitesimal in a way that puts the least constraints on fineness.
+
+\begin{defn}\label{infinitesimals}
+	Let $\stmt[u],\stmt[s] \in \tdomain$.
+Then $\stmt$ is \textbf{infinitesimal} with respect to $\stmt[u]$ if for any natural number $n$ we can find a set of $n$ statements $\{ \stmt_i \}_{i=1}^n$ such that:
+	\begin{enumerate}
+		\item $\stmt \finer \stmt_i$ for all $i$
+		\item $\stmt_i \ncomp \stmt_j$ for all $i \neq j$
+		\item $\bigOR\limits_{i=1}^n \stmt_i \finer \stmt[u]$ 
+	\end{enumerate}
+	In words, $\stmt$ is infinitesimal with respect to $\stmt[u]$ if for any number, there is a disjunction of that many broader mutually incompatible statements $\stmt_i$ is finer than $\stmt[u]$.
+\end{defn}
+
+\begin{remark}
+	Note that we do not require the $\stmt_i$ to be equigranular to $\stmt$.
+Also note that, with this definition, a statement can fail to be infinitesimal simply because one cannot find enough statements comparable to both $\stmt$ and $\stmt[u]$.
+\end{remark}
+
+\begin{coro}
+	The impossibility $\impossibility$ is an infinitesimal with respect to any $\stmt[u] \in \tdomain$.
+\end{coro}
+
+\begin{proof}
+	Let $\stmt[u] \in \tdomain$ and consider $\{ \stmt_i \}_{i=1}^n$ where $\stmt_i \equiv \impossibility$ for all $i$.
+Then $\impossibility \finer \impossibility \equiv \stmt_i$ for all $i$.
+Since $\impossibility \ncomp \impossibility$, $\stmt_i \ncomp \stmt_j$ for all $i$ and $j$.
+Lastly, $\bigOR\limits_{i=1}^n \stmt_i \equiv \impossibility \finer \stmt[u]$.
+This means that $\{ \stmt_i \}_{i=1}^n$ realizes infinitesimality for $\impossibility$, which is an infinitesimal of $\stmt[u]$.
+\end{proof}
+
+\begin{prop}\label{infinitesimal_conjunction}
+	Let $\stmt \in \tdomain$ be an infinitesimal with respect to $\stmt[u] \in \tdomain$. Then any $\stmt[v] \finer \stmt$ is also an infinitesimal with respect to $\stmt[u]$.
+\end{prop}
+\begin{proof}
+	Since $\stmt$ is an infinitesimal with respect to $\stmt[u]$, for any $n$ we can find a set of statements $\{ \stmt_i \}_{i=1}^n$ that realizes its infinitesimality.
+We have $\stmt[v] \finer \stmt \finer \stmt_i$ for all $i$.
+The other two properties do not depend on $\stmt$.
+Therefore $\{ \stmt_i \}_{i=1}^n$ realizes infinitesimality for $\stmt[v]$ as well so $\stmt[v]$ is an infinitesimal with respect to $\stmt[u]$.
+\end{proof}
+
+\begin{coro}
+	The conjunction of two infinitesimals with respect to $\stmt[u]$ is infinitesimal with respect to $\stmt[u]$.
+\end{coro}
+\begin{proof}
+	Let $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ be infinitesimal with respect to $\stmt[u]$.
+Since $\stmt[v]_1 \AND \stmt[v]_2 \narrower \stmt[v]_1$, $\stmt[v]_1 \AND \stmt[v]_2 \finer \stmt[v]_1$ and therefore $\stmt[v]_1 \AND \stmt[v]_2$ by \ref{infinitesimal_conjunction}.
+\end{proof}
+
+\begin{prop}
+	The disjunction of two infinitesimals with respect to $\stmt[u]$ that are comparable\footnote{Is comparability required?
+Can it be proved one way or another?} to each other is infinitesimal with respect to $\stmt[u]$.
+\end{prop}
+\begin{proof}
+	Let $\stmt[v],\stmt[w]$ be infinitesimal with respect to $\stmt[u]$.
+Without loss of generality, suppose $\stmt[v] \finer \stmt[w]$.
+Let $n>0$.
+Let $\{\stmt[w]_i\}_{i=1}^{2n}$ be a collection realizing infinitesimality for $2n$ for $\stmt[w]$.
+Define $\stmt_j = \stmt[w]_j\OR\stmt[w]_{j+n}$ for $j=1,\ldots,n$.
+Notice that the $\stmt_j$'s are mutually incompatible, and their disjunction must be finer than $\stmt[u]$.
+Further, since $\stmt[v]\finer\stmt[w]$ we have that $\stmt[v]\finer\stmt[w]_i$ for all $i$.
+It follows that:
+	$$
+	\stmt[v]\OR\stmt[w] \finer \stmt[w]_j\OR\stmt[w]_{j+n} = \stmt_j
+	$$
+	for all $j=1,\ldots,n$.
+Thus the collection $\{\stmt_j\}_{j=1}^n$ realizes $\stmt[v]\OR\stmt[w]$ being infinitesimal with respect to $\stmt[u]$ for $n$.
+We conclude that $\stmt[v]\OR\stmt[w]$ is infinitesimal with respect to $\stmt[u]$.
+\end{proof}
+
+\begin{prop}[Infinitesimality is consistent with fineness]\label{infinitesimal_finess}
+	Let $\stmt[u],\stmt[v]$ be statements such that $\stmt[v]\finer\stmt[u]$. Let $\stmt$ be an infinitesimal with respect to $\stmt[v]$. Then $\stmt$ is infinitesimal with respect to $\stmt[u]$. 
+\end{prop}
+\begin{proof}
+	Let $n>0$ be a natural number, and let $\{ \stmt_i \}_{i=1}^n$ be as in the definition of infinitesimal for $\stmt$ with respect to $\stmt[v]$.
+Then $\{ \stmt_i \}_{i=1}^n$ also satisfies the requirements for infinitesimal with respect to $\stmt[u]$.
+The first two requirements are unchanged, and the third is easy to see:
+	$$
+	\bigOR\limits_{i=1}^n \stmt_i \finer\stmt[v]\finer \stmt[u]
+	$$
+\end{proof}
+
+\begin{coro}[Infinitesimality is consistent with narrowness]
+	Let $\stmt[u],\stmt[v]$ be statements such that $\stmt[v]\narrower\stmt[u]$. Let $\stmt$ be an infinitesimal with respect to $\stmt[v]$. Then $\stmt$ is infinitesimal with respect to $\stmt[u]$.
+\end{coro}
+\begin{proof}
+	Since $\stmt[v]\narrower\stmt[u]$ implies $\stmt[v]\finer\stmt[u]$, proposition \ref{infinitesimal_finess} applies.
+\end{proof}
+
+We wish to also show some form of compatibility in the reverse direction.
+That is, if we shrink by a ``finite" amount, the sets which were previously infinitesimal should remain so (of course, if we shrink by an ``infinite" amount, e.g.
+by dropping a dimension, previously infinitesimal sets may no longer be so).
+
+\begin{prop}
+	Let $\tdomain$ be uniform.
+Let $\stmt[u],\stmt[v] \in \tdomain$ be statements such that $\stmt[v]$ is not infinitesimal with respect to $\stmt[u]$.
+Let $\stmt$ be a statement $\stmt\narrower\stmt[v]$ such that $\stmt$ is infinitesimal with respect to $\stmt[u]$.
+Then $\stmt$ is infinitesimal with respect to $\stmt[v]$.
+\end{prop}
+\begin{proof}
+	Note: the use of totality is highlighted as we would like to, if possible, remove it.
+
+	Let $\stmt[u]$, $\stmt[v]$, and $\stmt$ be as in the Proposition statement.
+We will show that $\stmt$ is infinitesimal with respect to $\stmt[v]$.
+To do so, we will show that for any $n>0$, we can find a set of $n$ statements $\{\stmt_i\}_{i=1}^n$ such that:
+	\begin{enumerate}
+		\item $\stmt \finer \stmt_i$ for all $i$
+		\item $\stmt_i \ncomp \stmt_j$ for all $i \neq j$
+		\item $\bigOR\limits_{i=1}^n \stmt_i \finer \stmt[v]$ 
+	\end{enumerate}
+	We will show this for arbitrary $n$, which will be sufficient to prove that $\stmt$ is infinitesimal with respect to $\stmt[v]$.
+
+	First, since $\stmt[v]$ is not infinitesimal with respect to $\stmt[u]$, there exists some $N>0$ such that for any set of $N$ statements, at least one of the conditions of infinitesimality of $\stmt[v]$ with respect to $\stmt[u]$ fails.
+
+	Let $n>0$.
+Let $\{\stmt_i\}_{i=1}^{nN}$ be a set of $n\cdot N$ statements realizing infinitesimality of $\stmt$ with respect to $\stmt[u]$ (i.e.
+they are mutually incompatible, broader than $\stmt$, and their disjunction is narrower than $\stmt[u]$).
+
+	{\bf By totality of $\finer$}\footnote{What we needed here is that there is a partition of the collection into $N$ sets of size $n$ such that for one of them, their disjunction is finer than each of $N-1$ disjunctions of the other subsets in the partition.}, we may assume the collection $\{\stmt_i\}_{i=1}^{nN}$ is ordered by fineness ($\stmt_1\finer\stmt_2\finer\cdots$).
+Consider the finest $n$ statements, that is, consider the collection $\{\stmt_i\}_{i=1}^{n}$.
+We will show that this collection satisfies the conditions for $\stmt$ being infinitesimal with respect to $\stmt[v]$ for $n$ statements.
+
+	By construction, this satisfies properties 1 and 2 for infinitesimality of $\stmt$ in $\stmt[v]$, since it is a subset of a collection which already satisfied those properties.
+
+	Now suppose it fails property 3\footnote{We will use totality since we need $\NOT \big(\stmt[x]\sfiner\stmt[y]\big) \implies \stmt[x]\coarser\stmt[y]$.}, that is, suppose 
+	$$
+	\bigOR\limits_{i=1}^n \stmt_i \scoarser\stmt[v].
+	$$
+	We will show that this will lead to a contradiction by implying that there is a collection of statements of size $N$ realizing $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$.
+
+	Consider now the collection of $N$ statements $\{\stmt[a]_j\}_{j=1}^N$ defined by $$\stmt[a]_j = \bigOR\limits_{i=n(j-1)+1}^{nj}\stmt_i.$$
+	These statements are built from the collection $\{\stmt_i\}_{i=1}^{n}$ by taking $N$ disjunctions each consisting of $n$ of the statements.
+Notice that $\stmt[a]_1$ is the disjunction of the collection $\{\stmt_i\}_{i=1}^{n}$ considered above.
+
+	First, this collection is mutually incompatible, since it is made up of disjunctions of mutually incompatible statements.
+This is property 2 for $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$.
+
+	Next, these must satisfy property 1 for $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$.
+To see why, notice that by our assumption, we have $\stmt[a]_1\scoarser\stmt[v]$, and by construction, we have $\stmt[a]_j\coarser\stmt[a]_1$ for all $j=1,\ldots,N$, and so $\stmt[v]\finer\stmt[a]_j$ for all $j=1,\ldots,N$.
+
+	Finally, they also must satisfy property 3: $$\stmt[a]_1\OR\cdots\OR\stmt[a]_N = \bigOR\limits_{i=1}^{nN}\stmt_i\finer\stmt[u].$$
+
+	Thus, this new collection violates the assumption that no set of $N$ statements satisfies all three properties for $\stmt[v]$ being infinitesimal in $\stmt[u]$.
+
+	{\bf By totality of $\finer$},\footnote{Here we are using totality as specified in the previous footnote.} we conclude that we must have $\bigOR\limits_{i=1}^n \stmt_i \finer\stmt[v]$ as required, and so the collection $\{\stmt_i\}_{i=1}^{n}$ realizes $\stmt$ being infinitesimal with respect to $\stmt[v]$ for $n$ statements.
+Since $n$ was arbitrary, it must be the case that $\stmt$ is infinitesimal with respect to $\stmt[v]$.
+\end{proof}
+
+\section{Notes}
+
+\subsection{Mark Notes again again}
+
+Question: Fix $\stmt[u]\in\tdomain$, and assume the following holds.
+Let $\stmt\in\tdomain$ such that $\stmt$ is compatible with infinitely many possibilities.
+Then there exist statements $\stmt_1,\stmt_2$ with $\stmt_1\ncomp\stmt_2$ and $\stmt_1\eqgran\stmt_2$ such that $\stmt_1\OR\stmt_2\narrower\stmt$ and $\stmt\AND\NOT(\stmt_1\OR\stmt_2)$ is an infinitesimal with respect to $\stmt[u]$.
+This means we can split statements in half (up to infinitesimals, with respect to fineness).
+Further, assume $\stmt[u]$ does not admit countable infinitesimals. That is, for any countable collection $S\subseteq\tdomain$ of infinitesimals, $\OR S$ is still an infinitesimal.
+
+We wish to know whether $\leq_{\stmt[u]}$ will have monotone continuity up to infinitesimals (this may or may not be related to the above).
+While we cannot quite answer that yet since we have not proven that $\lequ$ is transitive, we can think about it for $\finer$.
+The statement to understand is:
+
+Let $\{\stmt_i:i\in\mathbb{N}\}$ be a sequence of statements monotone increasing in narrowness (i.e.
+$\stmt_i\narrower\stmt_j$ for all $i\leq j$), and write $\stmt=\OR_i\stmt_i$ be the disjunction.
+Assume that $\stmt$ is finitely comparable to $\stmt[u]$.
+If $\stmt[t]\in\tdomain$ such that $\stmt[t]$ is finitely comparable to $\stmt[u]$, and $\stmt_i\finer\stmt[t]$ for all $i$, then $\stmt\finer\stmt[t]\OR\stmt[w]$ for some infinitesimal (with respect to $\stmt[u]$) $\stmt[w]$.
+
+Idea: added terms approach infinitesimal.
+
+Consider the new sequence defined by $\stmt'_i := \stmt_i \AND \NOT\stmt_{i-1}$ for $i\geq2$ and $\stmt'_1 := \stmt_1$.
+Then $\OR_{i=1}^n\stmt'_i = \stmt_n\finer\stmt[t]$ for all $n$.
+It follows that $\stmt = \OR_{i=1}^{\infty}\stmt'_i$, and so $\stmt$ is a countably infinite disjunction of disjoint statements.
+Since $\stmt$ is finitely comparable to $\stmt[u]$, not all of the $\stmt_i'$s can be infinitesimals.
+
+Consider $\stmt[v]_n = \OR_{i=n}^{\infty}\stmt'_i = \stmt \AND\NOT\stmt_{n-1}$.
+Observe that $$\AND_{n=1}^{\infty}\stmt[v]_n = \AND_{n=1}^{\infty}\stmt\AND\NOT\stmt_{n-1} = \stmt\AND\NOT\OR_{n=1}^{\infty}\stmt_n = \stmt\AND\NOT\stmt \equiv \impossibility$$
+
+We wish to show that $\stmt = \stmt_{n-1}\OR\stmt[v]_n\finer \stmt[t]\OR \stmt[w]$.
+Or equivalently, assuming there exist statements $\stmt[x]_n\eqgran\stmt_n$ with $\stmt[x]_n\narrower\stmt[t]$ and $\stmt[x]_{n-1}\narrower\stmt[x]_n$, we wish to show that $\stmt[w]' := \stmt[t]\AND\NOT\stmt[x]$ is an infinitesimal, where $\stmt[x]=\OR_{i=1}^n\stmt[x]_i$.
+Note: if we have a very robust collection of possibilities and a dense enough fineness relation, we can ``move statements around" to find equigranular copies inside other statements, so the claim to show reduces to the case in this paragraph.
+If $\stmt[x]$ is not narrower than $\stmt[t]$, then there must exist some possibility in $\stmt[x]$ outside of $\stmt[t]$, and this must have shown up in some $\stmt[x]_i$.
+Is this sufficient?
+
+Suppose $\{\stmt_i\}_{i=1}^{\infty}$ is such that $\AND_{i=1}^{\infty}\stmt_i = \impossibility$?
+Not sure.
+We need more theorems about characterizing infinitesimals beyond the definition.
+We don't have much in the way of finding arbitrarily large (finite) sets of equigranular statements.
+
+Comment: Let $\stmt[u] = \mathbb{R}$.
+Then let $\stmt[t] = (0,\infty)$ and $\stmt = \mathbb{R}$.
+These are both finitely comparable to $\stmt[u]$.
+Next, let $\stmt_n = (-n,n)$.
+Then this is an increasing sequence of statements in narrowness, each of which is finer than $\stmt[t]$.
+But their disjunction is $\stmt$ which is (strictly) coarser than $\stmt[t]$, and no infinitesimal $\stmt[w]$ (with respect to $\stmt[u]$) will be enough to have $\stmt\finer\stmt[t]\OR \stmt[w]$.
+Interestingly, in this case, each $\stmt_n\AND\NOT\stmt_{n-1}$ is a countable infinitesimal in $\stmt[u]$.
+This shows us that a natural space which we might like to consider (e.g.
+$\mathbb{R}$) will fail the above conclusion; however, each set $(-n,n)$ for $n\geq1$ is a countable infinitesimal in $\mathbb{R}$, and so not admitting countable infinitesimals will be an important and strict condition.
+
+Let $\stmt[u]$ be a statement.
+Suppose there exist statements $\stmt_i$, $i=1,\ldots$ which are mutually incompatible, equigranular, and $\stmt[u] = \OR_{i=1}^{\infty}\stmt_i$.
+Then all the $\stmt_i$'s are countable infinitesimals.
+So if it is possible to split a statement into countably infinitely many incompatible equigranular statements, then it admits countable infinitesimals.
+Not admitting countable infinitesimals is a strict one, but examples are provided by the standard Borel measure on bounded intervals in $\mathbb{R}$.
+
+\subsection{Unique measures}
+
+One condition for the measure to be unique given a unit statements is that all points (i.e.
+possibilities in our framework) are equigranular.
+In fact, consider a domain with just three points not equigranular to each other, but each finer than the other.
+We can construct any measure that satisfies the requirement by picking three numbers and assigning them accordingly, since we can't determine ``how bigger'' is one point with respect to the other.
+
+\subsection{Finding a probability measure}
+
+Our first goal is to find a measure which is compatible with the fineness preorder.
+The idea is that fineness tells us which sets are ``larger" (less precise) than each other, regardless of any overlap.
+A measure will quantify this and allow us to assign numbers to the sizes.
+
+The paper \cite{probexistence} seems to be extremely relevant here.
+The basic idea in this paper is to find necessary and sufficient conditions for the existence of a probability measure on a Boolean algebra (e.g.
+of subsets of some set) which is compatible with a total pre-order.
+There are a few points to keep in mind.
+
+First, we need a \emph{total} pre-order.
+This means for every pair of elements, one is finer than the other.
+This is certainly not the case for fineness in general: color and position, for example, cannot be compared.
+This is the goal of defining the \emph{unit} above.
+The unit serves as a baseline statement against which others can be compared for fineness.
+For example, a statement about position singles out length as the dimension to compare, or a statement about numbers of objects specifies ``count" as the dimension to compare (e.g.
+the discrete unit of number of objects).
+
+Next, we need the three properties specified in the paper, to be written up later! We are optimistic that these three properties are present in our situation.
+
+There is one additional caveat.
+The measure constructed by the main theorem in the above-mentioned paper is always a probability measure.
+This means the whole space has finite measure.
+For example, one could put the Lebesgue measure on $(0,1)$ but under a homeomorphism $(0,1)\to\mathbb{R}$ induce a finite (probability) measure on  $\mathbb{R}$.
+If we chose the full set as a unit, we are exactly looking for a probability measure.
+If we pick a unit that is not the full set, we can also "constrain" the problem to that subset, so that we still have a probability measure.
+Then we may be able to "stich together" the measures constructed on all equigranular sets assuming they cover the full space.
+
+\subsection{Applying results of the paper \cite{villegas}}
+
+In this part, we will assume that $\tdomain_{\stmt[u]}$ is a subset of the theoretical domain, $\stmt[u]\in \tdomain_{\stmt[u]}$ is a distinguished element (we'll refer to it as the \emph{unit} when needed), and there is a total pre-order $\leq_{\stmt[u]}$ defined on $\tdomain_{\stmt[u]}$.
+We will write $=_{\stmt[u]}$ for $\leq_{\stmt[u]}$ and $\geq_{\stmt[u]}$, and we write $<_{\stmt[u]}$ for $\leq_{\stmt[u]}$ and not $=_{\stmt[u]}$.
+We will use the word ``equiprobable" for $=_{\stmt[u]}$ since it means something different from equigranular (since $\leq_{\stmt[u]}$ is a different ordering from fineness, even though it will be derived from fineness).
+We assume $\tdomain_{\stmt[u]}$ is closed under complement and countable union (and hence countable intersection), and that there are two additional elements, $\certainty$ and $\impossibility$, which are maximal and minimal in this ordering, respectively.
+
+\begin{remark}
+In defining the subset $\tdomain_{\stmt[u]}$ and the total preorder $\leq_{\stmt[u]}$, we will have to be very careful.
+We want to ``squash" together all the sets which should be measure zero in this unit (e.g.
+anything ``lower-dimensional"), ``blow up" all the ``infinite measure" subsets (e.g.
+``higher dimensional"), and ensure that everything is compatible under this order, so there can be only one ``type" of property under consideration.
+\end{remark}
+
+We write out the main axioms and statements from \cite{villegas} which interest us in this situation, in the notation established earlier. On page 1789, we have the following:
+
+Axiom of Pre-order:
+\begin{itemize}
+    \item Totality: If $\stmt_1,\stmt_2\in \tdomain_{\stmt[u]}$, then $\stmt_1 \leq_{\stmt[u]} \stmt_2$ or $\stmt_2\leq_{\stmt[u]} \stmt_1$
+    \item Reflexivity: For all statements $\stmt\in \tdomain_{\stmt[u]}$, we have $\stmt\leq_{\stmt[u]}\stmt$
+    \item Transitivity: For all statements $\stmt_1,\stmt_2,\stmt_3\in \tdomain_{\stmt[u]}$, if $\stmt_1\leq_{\stmt[u]}\stmt_2$ and $\stmt_2\leq_{\stmt[u]}\stmt_3$, then $\stmt_1\leq_{\stmt[u]}\stmt_3$.
+    \item Existence of max and min: We have $\impossibility\leq_{\stmt[u]}\certainty$ and \emph{not} $\certainty\leq_{\stmt[u]}\impossibility$.
+For all $\stmt\in \tdomain_{\stmt[u]}$ we have $\impossibility\leq_{\stmt[u]}\stmt\leq_{\stmt[u]}\certainty$
+\end{itemize}
+
+The above list is simply the requirement that $\leq_{\stmt[u]}$ is a total pre-order and that there are a maximal and minimal element.
+
+Axiom of Monotony: 
+Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
+Assume that $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, and $\stmt_2\ncomp \stmt_4$ (i.e.
+$\stmt_2\AND \stmt_4 =_{\stmt[u]} \impossibility$).
+Then we have $\stmt_1\OR \stmt_3 \leq_{\stmt[u]} \stmt_2\OR \stmt_4.$
+
+Monotony guarantees that given two incompatible statements, each ``bigger" than another statement, then their disjunction is still bigger than the disjunction of the smaller statements.
+
+The following are consequences of the two axioms above. Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
+\begin{itemize}
+    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ and $\stmt_2<_{\stmt[u]} \stmt_3$, then $\stmt_1<_{\stmt[u]}\stmt_3$. 
+    \item If $\stmt_1\narrower \stmt_2$ then $\stmt_1\leq_{\stmt[u]} \stmt_2$
+    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, $\stmt_2\leq_{\stmt[u]} \stmt_4$, and $v\leq_{\stmt[u]} \stmt_1$, then $\stmt_2 \AND \NOT \stmt_1 \leq_{\stmt[u]} \stmt_4 \AND \NOT \stmt_3$
+    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ then $\NOT \stmt_2 \leq_{\stmt[u]}\NOT \stmt_1$
+\end{itemize}
+
+Next, we will need a more subtle axiom in order to make sense of infinite combinations of statements.
+
+\begin{remark}
+The \emph{fineness} relation $\finer$ should satisfy all of the axioms of pre-order and the axiom of monotony, except for totality. We restrict to $\tdomain_{\stmt[u]}$ in order to get totality. 
+\end{remark}
+
+Axiom of Monotone Continuity:
+Let $\{\stmt_i : i\in \mathbb{Z}^+\}$ be a sequence of statements which is monotone increasing with respect to narrowness (containment), and let $\stmt = \cup_i\stmt_i$ be the disjunction.
+Let $t\in \tdomain_{\stmt[u]}$ such that $\stmt_i\leq_{\stmt[u]} t$ for all $i$.
+Then $\stmt\leq_{\stmt[u]} t$.
+
+In other words, an infinite disjunction can only strictly exceed a statement in coarseness if the disjunction of a finite subset of them already exceeds it.
+Here is an interesting consequence of the combination of the above axioms:
+
+\begin{prop}[Corollary to Lemma 1 in \cite{villegas}]
+\label{infinitenull}
+Let $\{\stmt_i : i\in \mathbb{Z}^+\}$ be an infinite collection such that $\stmt_i\AND\stmt_j =_{\stmt[u]} \impossibility$ for all $i\neq j$, and $\stmt_i =_{\stmt[u]} \stmt_j$ for all $i,j$.
+Then it must be the case that $\stmt_i=_{\stmt[u]}\impossibility$ for all $i$.
+\end{prop}
+This says that infinite collections of equiprobable statements with null overlaps must be null themselves.
+
+An \emph{atom} is a statement $\stmt\in S$ such that $\impossibility<_{\stmt[u]}\stmt$ and there does not exist any statement $\stmt_2\in S$ such that $\impossibility<_{\stmt[u]} \stmt_2<_{\stmt[u]} \stmt$.
+In particular, Theorem 5 in \cite{villegas} \S3 includes the fact that $\tdomain_{\stmt[u]}$ is atomless if and only if every statement can be partitioned into two equiprobable statements.
+Theorem 3 in \S4 states that if $\tdomain_{\stmt[u]}$ is atomless, then there exists a unique probability measure compatible with $\leq_{\stmt[u]}$, and it is countably additive!
+
+\subsection{Using Monotone Continuity}
+
+The core of the proof of Theorem 3 in Section 4 of \cite{villegas} comes from Theorem 5 (iv) in Section 3.
+In the terminology of Villegas, a \emph{qualitative probability $\sigma$-algebra} is a $\sigma$-algebra of subsets of a space with an ordering that satisfies the Axiom of Pre-order, the Axiom of Monotony, and the Axiom of Monotone Continuity.
+We also need the language of \emph{atoms}, defined earlier.
+We will reproduce some of the ideas in the proofs of these theorems in the language of statements and our theoretical domain, while attempting to keep track of the usages of Monotone Continuity.
+The hope is that there is a weakened version of the axiom that gets us the desired result, or that we can build the ordering $\leq_{\stmt[u]}$ with Monotone Continuity but it can still distinguish the infinite sets described earlier.
+
+Let $A$ be a statement in $\tdomain_{\stmt[u]}$ with $A>_{\stmt[u]} \impossibility$. A \emph{minor incomplete partition} of $A$ is a collection $M_A$ of statements with the following properties:
+\begin{itemize}
+    \item We have $\stmt>_{\stmt[u]}\impossibility$ and $\stmt \narrower A$ for all $\stmt\in M_A$, that is, they are non-null and narrower than $A$
+    \item For all $\stmt_1,\stmt_2\in M_A$, we have $\stmt_1\AND\stmt_2 =_{\stmt[u]} \impossibility$, that is, they overlap only on statements ``squashed" to zero. 
+    \item The disjunction of $M_A$ is finer than its complement in $A$: $$\bigOR_{\stmt\in M_A}\stmt \finer (A\AND \NOT\bigOR_{\stmt\in M_A}\stmt).$$ This is the condition of being \emph{minor}.
+    %UM_A finer than A - UM_A^C
+\end{itemize}
+
+It is shown in \cite{villegas} that any (minor) incomplete partition is at most countable (Theorem 3).
+Rather than give the full proof of this fact, we note that it follows from a large collection of lemmas, the most interesting of which is the following, which we will prove below:
+
+\begin{prop}[Lemma 1 from \cite{villegas}]
+If $\{\stmt[a]_i: i\in I\}$ is an infinite set of statements such that (1) $\stmt[a]_i \AND \stmt[a]_j =_{\stmt[u]}\impossibility$ for all $i\neq j$ and (2) $\stmt[a]_i>_{\stmt[u]}\impossibility$ for all $i$, then the following holds.
+Given a statement $\stmt>_{\stmt[u]}\impossibility$, only finitely many of the $\stmt[a]_i$'s have $\stmt[a]_i>_{\stmt[u]}\stmt$.
+\end{prop}
+\begin{proof}
+By way of contradiction, let $\stmt[a]_1,\ldots$ be a sequence of statements violating the conclusion (i.e.
+infinitely many statements $\stmt[a]_i$ with $\stmt[a]_i>_{\stmt[u]}\stmt>_{\stmt[u]}\impossibility$.
+Consider the new sequence $$\stmt_i = \bigOR\limits_{n=i}^{\infty}\stmt[a]_n.$$ This sequence is monotonely decreasing to the impossibility in narrowness (and hence in fineness and in $>_{\stmt[u]}$), but we also have $\stmt_i\geq_{\stmt[u]}\stmt$ for all $i$.
+By monotone continuity (the contrapositive), we have $\stmt[s]\leq_{\stmt[u]}\impossibility$, a contradiction.
+\end{proof}
+
+A quick corollary of this is Proposition \ref{infinitenull}, above. Using minor incomplete partitions, we prove the following.
+
+\begin{prop}[Theorem 4, \S3 \cite{villegas}]
+\label{split}
+If there are no atoms, then for every statement $\stmt$ there exist statements $\stmt_1$ and $\stmt_2$ such that: $\stmt_1\ncomp \stmt_2$, $\stmt_1\equ\stmt_2$, and $\stmt_1\OR\stmt_2 = \stmt$.
+In other words, $\stmt$ can be partitioned into two equiprobable (note: I remember we don't like this word but I don't remember if we replaced it or not) statements.
+\end{prop}
+\begin{proof}
+Let $S$ be the set of minor incomplete partitions of $\stmt$, and endow it with a partial order by inclusion.
+Given a totally ordered chain in $S$, its union is still a minor incomplete partition, and is thus an upper bound for the chain.
+We may apply Zorn's lemma to find a maximal element of $S$, the disjunction of which we call $\stmt_1$.
+We claim that $\stmt_1$ and $\stmt_2:= \stmt\AND\NOT\stmt_1$ are equiprobable, incompatible statements.
+Incompatibility is immediate from the definition.
+For equiprobability, first note that by definition of being a minor incomplete partition we have $\stmt_1\lequ\stmt_2$.
+Next, if $\stmt[t]_1,\ldots$ is a monotone decreasing (with respect to $\lequ$) sequence of statements $\stmt[t]_n>_{\stmt[u]}\impossibility$ narrower than $\stmt_2$ and converging to a zero-comparable event, we have that $\stmt_1\OR\stmt[t]_n>_{\stmt[u]}\stmt_2\OR\stmt[t]_n$ since $\stmt_1$ is maximal.
+We needed no atoms in order to guarantee this sequence exists (Lemma 3 in Section 3 of \cite{villegas}).
+It follows by monotone continuity that $\stmt_1\geq_{\stmt[u]} \stmt_2$.
+\end{proof}
+
+We will also need a few types of \emph{random variables} as defined in \cite{villegas} \S3.
+Let $A({\stmt[u]}) = \{x \in X : x \comp \stmt[u]\}$ be the set of possibilities associated with $\stmt[u]$.
+This is a subset of the collection of possibilities which will serve as the ``sample space" after picking a unit $\stmt[u]$.
+
+A \emph{uniform sequence of finitely-valued random variables} is a sequence of functions (random variables in the traditional sense) $X_1,X_2,\ldots: A({\stmt[u]})\to E$ where $E$ is a finite set, with the following property: for all $\{x_1,\ldots,x_n\}\subseteq E$, all the events $$\{p\in A({\stmt[u]}): X_i(p) = x_i, i=1,\ldots,n\}$$
+are equiprobable ($\equ$ with each other).
+
+Notice that monotone continuity allows us to let $n\to\infty$ to define events where the whole sequence is specified with the resulting events all being equiprobable.
+By Proposition \ref{infinitenull} above, all such events are null.
+
+Building on these sequences of random variables, we will use \emph{uniform random variables} to define probability measures.
+Let $X:A({\stmt[u]})\to[0,1]$ be a random variable, and let $I$ and $J$ be intervals contained in $[0,1]$.
+We say $X$ is a uniform random variable if $X^{-1}(I)\lequ X^{-1}(J)$ if and only if $|I|\leq|J|$ for all such $I,J$.
+
+From Proposition \ref{infinitenull}, $X^{-1}(a)$ is a zero-comparable event for all $a\in[0,1]$, and if $0=a_0<a_1<\cdots<a_n=1$, then $X^{-1}([a_i,a_{i+1}))$ are a finite partition of $\tdomain$.
+
+Next, we will pull many of the above concepts together to show how we obtain a probability measure on $A({\stmt[u]})$ realized as a function on $\tdomain$ compatible with $\lequ$.
+
+\begin{prop}\label{uniformexist}
+If $\tdomain_{\stmt[u]}$ has no atoms with respect to $\lequ$, then there exists a uniform random variable.
+\end{prop}
+\begin{proof}
+By Proposition \ref{split} we may split any statement into two equiprobable, null-compatible sets.
+Starting with the statement $\stmt[u]$ (here, the ``universe" which we want to have probability 1), repeatedly divide into two statements in stages.
+At stage $n$ (starting with $n=0$), we will have statements $\stmt[u]^{(n)}_1,\ldots,\stmt[u]^{(n)}_{2^n}$.
+Now, take the disjunction of alternating statements in each collection so that there is a ``right" and ``left" (odd and even index) half at each stage.
+For example, at stage $n=1$, the odd and even halves are simply $\stmt[u]^{(1)}_1$ and $\stmt[u]^{(1)}_2$; at stage $n=2$ we have $\stmt[u]^{(2)}_1\OR\stmt[u]^{(2)}_3$ and $\stmt[u]^{(2)}_2\OR\stmt[u]^{(2)}_4$.
+
+Define a sequence of random variables $X_n:A(\stmt[u])\to\{0,1\}$ by $X_n(a) = 1$ if and only if $a$ is in the odd half in stage $n$.
+This gives each $a\in A(\stmt[u])$ a binary ``address".
+By the construction, it follows that this is a uniform sequence of finitely-valued random variables (defined above).
+
+Now, we define a uniform random variable $X:A(\stmt[u])\to[0,1]$.
+For $a\in A(\stmt[u])$, define $X(a) = \sum_{n=1}^{\infty}2^{-n}X_n(a)$.
+In other words, $X(a)$ is the binary expansion $0.X_1(a)X_2(a)\ldots_2$.
+
+One can show immediately that $X^{-1}(I)\lequ X^{-1}(J)$ if and only if $|I|\leq |J|$ when $I$ and $J$ have endpoints on numbers with finite binary expansions (since $X_1,\ldots,X_n$ are uniform), and by monotone continuity, the uniformity condition extends to arbitrary intervals in $[0,1]$.
+\end{proof}
+
+Finally, we reach our goal of building a probability measure on $A(\stmt[u])$.
+
+\begin{prop}
+If $\tdomain_{\stmt[u]}$ has no atoms with respect to $\lequ$, then there exists a unique compatible probability measure, and it is countably additive. 
+\end{prop}
+\begin{proof}
+By Propsition \ref{uniformexist}, let $X:POSS_{\stmt[u]}\to[0,1]$ be a uniform random variable.
+%Define $F:[0,1]\to\tdomain$ by $F(x) = X^{-1}([0,x])$.
+It is straightforward to see that $F$ is monotone in $x$ (with respect to $\lequ$), and that if $x_n\to x$, then $\lim_nF(x_n) \equ F(x)$.
+%Started writing this part from paper, but it is not directly used in construction.
+Will add details later as needed.
+We claim that for any statement $\stmt\in\tdomain_{\stmt[u]}$, there exists a unique number $p\in[0,1]$ such that $\stmt\equ X^{-1}([0,p])$.
+Define $P:\tdomain_{\stmt[u]}\to[0,1]$ by $P(\stmt)= p$.
+This defines a probability measure compatible with $\lequ$.
+Details forthcoming as needed....
+\end{proof}
+
+Ideas involved: Defines a function $X:A(\stmt[u])\to [0,1]$ in a way that is consistent with $\lequ$ by having ``larger" sets of possibilities (i.e.
+coarser statements) corresponding to larger intervals in $[0,1]$.
+Then, the measure of any other statement is given by finding an equiprobable statement which is the preimage of $[0,p]$ under $X$ and defining the measure to be $p$.
+
+Wish: find a way to prove the above statements in an order-theoretic way.
+In particular, we wish to show that the order type of the equigranular-equivalence classes is equivalent to (a subset of) $\mathbb{R}$.
+Sufficient conditions to be isomorphic to $\mathbb{R}$ are (1) lacking endpoints (2) complete (i.e.
+has Least Upper Bound property) (3) has dense countable subset.
+The use of random variables may be very similar to an approach using order theory.
+For example, the transition from a uniform sequence of finitely-valued random variables into a (single) uniform random variable makes use of a countable dense subset.
+
+\subsection{Example with lack of monotonicity}
+
+Consider a Euclidian two dimensional space charted by $\{x,y\}$.
+Let $U_{i,j}$ be the square in the interval $x \in (i, i+1)$ and $y \in (j, j+1)$.
+Let $V$ be the vertical (infinite) strip in the interval $x \in (0, 1)$ and $W$ the vertical strip in the interval $x \in (0, 2)$.
+Define the sequence $W_i = \{ U_{0,0}, U_{1,0}, U_{0,1}, U_{1,1}, U_{0,2}, U_{1,2}, U_{0,3}, U_{1,3}, ...
+\}$.
+We have $\bigcup\limits_0^\infty W_i = W$.
+This means $\bigcup\limits_0^n W\finer V$ for all finite $n$ but $\bigcup\limits_0^\infty W_i = W \coarser V$.
+
+\subsection{Some thoughts on Example in 3.3}
+
+The construction above hints at some fundamental concepts in measure theory which will be important for us to consider.
+
+Notice that the two sets $V$ and $W$ are both of infinite Lebesgue measure (area).
+With respect to Lebesgue measure, monotone continuity is immediate since if $\bigcup\limits_0^n W\finer V$ for all $n$ then $\lim\limits_{n\to\infty}\mu(\bigcup\limits_0^n W) = \mu(W)\finer V$.
+Even the above example does not violate this, since there, both would be ``equivalent" in fineness, namely they are infinitely coarse.
+
+Our goal is to have a measure which is able to distinguish sets which may be of infinite Lebesgue measure as well.
+This will be very difficult to fully capture.
+Consider the following procedure.
+Starting with the vertical strip $V$, slide all the squares at even $y$-values to the right by one, and then allow the squares to collapse down to rest on the square below.
+Now, by simply rearranging regions, we have (almost) a cover of the wider strip $W$.
+The only missing points are the line $x=1, y>0$.
+If instead we used $(0,1]$ and $(0,2]$ for the strips, this procedure would cover them exactly.
+One could continue and cover the entire plane with similar translations.
+
+Here is one approach to distinguish $V$ and $W$, and similar methods will enable one to distinguish many more infinite-measure subsets.
+Let $S_n$ be the $2n\times 2n$ square centered the origin with vertices at $(\pm n, \pm n)$.
+Then for all $n$, $V\cap S_n$ has area $n$, and $W\cap S_n$ has area $2n$.
+This holds for all $n$ and so we might use something like this to think about how we would distinguish infinite areas.
+Notice that this is in some sense writing $V$ and $W$ as countable unions of nested sets in a standardized way (e.g.
+$V = \bigcup\limits_n V\cap S_n$) and then comparing the finite-area subsets.
+One could choose shapes other than $S_n$ for this purpose.
+Another way to distinguish them would be to look at the limit $\lim\limits_{n\to\infty}\text{Area}(V\cap S_n)/\text{Area}(S_n)$; in this case they are both 0 but for a different choice of $S_n$ one could find an asymptotic difference.
+The point of the above discussion is to try to gain a better idea of which (infinite/unbounded) sets we want to be related by $\finer$.
+It is not clear even for ``nice" regions of $\mathbb{R}^2$, let alone at the full generality we are seeking.
+
+Regardless, the goal is not just to distinguish infinite-Lebesgue-measure sets. We need to understand when monotone continuity is relevant and what happens without it.
+
+\subsection{Infinitesimals}
+
+\subsubsection{Equivalence at $\stmt[u]$ scale}
+
+\def\equivu{\cong_{\stmt[u]}}
+
+\begin{defn}
+Let $\stmt_1, \stmt_2 \in \tdomain$.
+We say $\stmt_1$ is \textbf{equivalent to $\stmt_2$ at $\stmt[u]$ scale}, noted $\stmt_1 \equivu \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$.
+\end{defn}
+
+\begin{prop}
+Equivalence at $\stmt[u]$ scale is an equivalence class.
+\end{prop}
+\begin{proof}
+For reflexivity, let $\stmt \in \tdomain$.
+We have $\stmt \OR \impossibility \equiv \stmt \OR \impossibility$.
+Since $\impossibility$ is an infinitesimal for all $\stmt[u]$ (TODO: add proposition), $\stmt \equivu \stmt$.
+For symmetry, note that the definition itself is symmetrical.
+For transitivity, let $\stmt_1 \equivu \stmt_2 \equivu \stmt_3$.
+Then we have $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$ and $\stmt_2 \OR \stmt[w]_2 \equiv \stmt_3 \OR \stmt[w]_3$.
+\end{proof}
+
+\subsubsection{Greatness in $\stmt[u]$}
+
+First define ``as big in $\stmt[u]$'' $\equ$.
+
+\begin{defn}
+Let $\stmt_1, \stmt_2 \in \tdomain$.
+Then $\stmt_1$ is \textbf{as big as} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \equ \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$.
+\end{defn}
+
+\begin{prop}
+Suppose $\tdomain$ is uniform.
+Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$, then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
+\end{prop}
+
+We start with a simpler case:
+
+\begin{prop}
+Let $\stmt_1,\stmt,\stmt[v]_1\in\tdomain$ such that $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$, and suppose $\stmt[v]_1$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt\equ\stmt_1$. 
+\end{prop}
+\begin{proof}[Using Prop \ref{uniformconsequence}]
+If both of the $\finer$'s are actually equigranular, there is nothing to show.
+Next consider the case that $\stmt_1\sfiner\stmt\finer\stmt_1\OR\stmt[v]_1$.
+Then by Proposition \ref{uniformconsequence} there exists $\stmt_2$ such that $\stmt_2\narrower\stmt$ and $\stmt_2\eqgran\stmt_1$.
+But how do we put this together??
+What about the even simpler version of the lemma with $\narrower$ instead of $\finer$?
+Just use $\stmt\OR (\stmt_1\OR\stmt[v]_1 \AND \NOT\stmt)$.
+\end{proof}
+
+This seems problematic already:
+
+We have $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$.
+We need to find some $\stmt[w],\stmt[w]_1$ infinitesimal with respect to $\stmt[u]$, such that $\stmt\OR\stmt[w]\eqgran\stmt_1\OR\stmt[w]_1$.
+Since we don't know anything about the overlap of $\stmt$ and $\stmt_1$, we can't use any operations like $\stmt\AND\NOT\stmt_1$ to find something small.
+We need to know that additional statements exist in order to find $\stmt[w]$ and $\stmt[w]_1$.
+
+For example, what if every infinitesimal statement was narrower $(\narrower)$ than $\stmt$?
+In this case, we could have something like $\stmt_1\sfiner\stmt\sfiner\stmt_1\OR\stmt[x]$ for all statements (including all infinitesimals) $\stmt[x]$.
+It would then not be the case that $\stmt\equ\stmt_1$.
+
+\begin{prop}[Need something like this but not clear if correct!]
+Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$. Then there exist statements  $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$, infinitesimal in $\stmt[u]$, such that:
+$$
+\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
+$$
+\end{prop}
+
+Note that the existence of the right combination of incompatible statements is sufficient.
+However, that is stronger than we need and may be harder to prove.
+Problem: we have very few propositions about existence of statements!
+It is hard to ensure certain ones exist.
+
+Consider the following (extreme) case. Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$, but the infinitesimals which realize these relationships are $\stmt[v]_1$ and $\stmt[v]_2$, as follows:
+$$
+\stmt_1\OR\stmt[v]_1 \eqgran \stmt_2\OR\stmt[v]_2,\ \stmt_2\OR\stmt[v]_1\eqgran\stmt_3\OR\stmt[v]_2
+$$
+Without further axioms, it may not be possible to put this together to get transitivity (i.e. to find that $\stmt_1\equ\stmt_3$).
+
+\begin{prop}
+The relation $\equ$ is an equivalence relation.
+\end{prop}
+\begin{proof}
+Reflexivity and symmetry are immediate since $\eqgran$ is an equivalence relation.
+For transitivity, suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$.
+Let $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$ be infinitesimals with respect to $\stmt[u]$, such that $\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2$ and $\stmt_2\OR\stmt[w]_2\eqgran\stmt_3\OR\stmt[w]_3$.
+Then we have:
+$$
+\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
+$$
+Since $\stmt[v]_1\OR\stmt[w]_2$ and $\stmt[w]_3\OR\stmt[v]_2$ are infinitesimal with respect to $\stmt[u]$ (being finite disjuncts of infinitesimals), we have $\stmt_1\equ\stmt_3$. 
+\end{proof}
+
+\begin{defn}\label{smalleru}
+Let $\stmt_1, \stmt_2 \in \tdomain$.
+Then $\stmt_1$ is \textbf{smaller than} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \lequ \stmt_2$, if $\stmt_1 \finer \stmt_2$ or $\stmt_1 \equ \stmt_2$.
+We define $\stmt_1\leu\stmt_2$ by $\stmt_1\lequ\stmt_2$ and not $\stmt_1\equ\stmt_2$.
+\end{defn}
+{\bf Warning:} Transitivity doesn't seem to work as stated (see the proof attempt below).
+We propose the following modification: say $\stmt_1\lequ\stmt_2$ if there exist statements $\stmt_1'$ and $\stmt_2'$ with $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ such that $\stmt_1'\finer\stmt_2'$.
+Note that {\bf by totality}, it is sufficient to just say $\stmt_1\finer\stmt_2'$ (no need for $\stmt_1'$).
+
+ \begin{prop}
+$\stmt_1 \leu \stmt_2$ if for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$. 
+\end{prop}
+\begin{proof}
+Since $\impossibility$ is infinitesimal, we have $\stmt_1 = \stmt_1\OR\impossibility \sfiner \stmt_2$, and so by Definition \ref{smalleru}, we have that $\stmt_1\lequ\stmt_2$.
+Now, we must show it is not the case that $\stmt_1\equ \stmt_2$.
+
+Let $\stmt[v]_1,\stmt[v]_2\in\tdomain$ be infinitesimals.
+Then $\stmt_1\OR\stmt[v]_1 \sfiner \stmt_2 \finer\stmt_2\OR\stmt[v]_2$.
+We conclude that it cannot be the case that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$ for any infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_1,\stmt[v]_2$.
+\end{proof}
+
+We next show that $\lequ$ is a pre-order, and that it is a partial order with respect to $\equ$-equivalence classes.
+
+\begin{prop}
+The relation $\lequ$ is reflexive and transitive.  
+\end{prop}
+\begin{proof}
+Reflexivity is immediate since for any $\stmt\in\tdomain$, we have $\stmt\finer\stmt$, so $\stmt\lequ\stmt$. (This works for both the new and old definition)
+
+Transitivity requires a little more work. Let $\stmt_1,\stmt_2,\stmt_3\in\tdomain$ with $\stmt_1\lequ\stmt_2\lequ\stmt_3$.
+
+{\bf Using the original definition:}
+We have several cases to consider.
+
+\emph{Case 1:} $\stmt_1\finer\stmt_2\finer\stmt_3$. Then $\stmt_1\finer\stmt_3$ and so $\stmt_1\lequ\stmt_3$.
+
+\emph{Case 2:} $\stmt_1\equ\stmt_2\equ\stmt_3$. Then $\stmt_1\equ\stmt_3$ (since $\equ$ is an equivalence relation), and so $\stmt_1\lequ\stmt_3$.
+
+\emph{Case 3:} $\stmt_1\equ\stmt_2\finer\stmt_3$. Then for some infinitesimals $\stmt[v]_1$ and $\stmt[v]_2$ (with respect to $\stmt[u]$), we have:
+$$
+\stmt_1\finer\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2\finer\stmt_3\OR\stmt[v]_2
+$$
+Unclear what to do here -- need $\stmt_3$ alone on the right-hand side somehow. Case 4 has the same problem. Let's try it with the modified definition (in the ``warning").
+
+{\bf Using the new definition:}
+Let $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ with $\stmt_1'\finer\stmt_2'$, and let $\stmt_2''\equ\stmt_2$ and $\stmt_3''\equ\stmt_3$ with $\stmt_2''\finer\stmt_3''$. 
+Then for an appropriate choice of infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_i,\stmt[w]_i,\stmt[v]_i',\stmt[w]_i'$, $i=1,2,3$, we have:
+$$
+\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_1'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \finer \stmt_2'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_2''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \finer
+$$
+$$
+\stmt_3''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \eqgran
+\stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1'
+$$
+Since $\stmt_1 \equ \stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3''$ and $\stmt_3 \equ \stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1' $, we have that $\stmt_1\lequ\stmt_3$.
+
+\end{proof}
+
+\begin{prop}
+If $\stmt_1\lequ\stmt_2$ and $\stmt_1\gequ \stmt_2$, then $\stmt_1\equ\stmt_2$.
+\end{prop}
+\begin{proof}
+With the original definition, this is immediate, since we would either have $\stmt_1\eqgran\stmt_2$ (which implies $\stmt_1\equ\stmt_2$), or we would already have $\stmt_1\equ\stmt_2$.
+
+With the modified definition, we need a bit more work. 
+\end{proof}
+
+Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotone?
+
+% case: possibilities in u: (1) finite: impose counting measure (2) countable: finitely additive but not countably (3) uncountable: countably additive, legitimate measure
+
+\subsection{Collapsing Orders}
+
+Searching around did not yield the theorems desired, so we will discuss the tools needed here.
+It turns out that it is not so straightforward to collapse a poset by an equivalence relation.
+Let's start with a possible definition:
+
+If $(P,\leq)$ is a poset and $\sim$ is a collapsing equivalence relation, then endow the set $P/\sim$ with the relation $\leq/\sim$ defined by $x/\sim \leq/\sim y/\sim$ whenever there exist $x'\in x/\sim$ and $y'\in y/\sim$ with $x'\leq y'$.
+
+It turns out that this is not sufficient. Here is an important example. Let our set be $\{w,x,y,z\}$ and a poset relation defined by $x\leq w$, $y\leq w$, $y\leq z$, with the following Hasse diagram:
+
+\begin{tikzpicture}
+    % First, locate each of the nodes and name them
+    \node (top) at (0,0) {$w$};
+    \node [below left  of=top] (left)  {$x$};
+    \node [below right of=top] (bottom right) {$y$};
+    \node [above right of=bottom right] (right) {$z$};
+
+    % Now draw the lines:
+    \draw [shorten <=-2pt, shorten >=-2pt] (top) -- (left);
+    \draw [shorten <=-2pt, shorten >=-2pt] (top) -- (bottom right);
+    \draw [shorten <=-2pt, shorten >=-2pt] (bottom right) -- (right);
+\end{tikzpicture}
+
+Let's consider two possible cases.
+Suppose first $w\sim y$, and $x$ and $z$ are only $\sim$ with themselves.
+Then in order to collapse $\leq$ under $\sim$, we will need $x \leq w\sim y \leq z$ for transitivity.
+
+Consider next the case that $w\sim y$ and $x\sim z$. In order to collapse this, we will need that $w\sim x\sim y \sim z$.
+
+These examples show that not every equivalence relation can collapse a poset into a poset (even if it only collapses a chain).
+
+Note that neither of the above problems are possible with {\bf totality}.
+It seems that what makes the most sense is collapsing by some equivalence relation, and then extending both the ordering and the equivalence relation to generate a legitimate poset.
+We can explore this later if needed.
+
+\section{Possible example}
+
+Let $\mathbb{N}$ be the set of natural numbers.
+We define the sets $D_{i,k} = \{ x \in \mathbb{N} \, | \, x \mod i = k \}$ with $k < i$.
+Therefore $D_{1,0} = \mathbb{N}$, $D_{2,0} = \{ 2, 4, 6, ...
+\}$, $D_{2,1} = \{ 1, 3, 5, ...
+\}$ and so on.
+Additionally, we say $D_{i,j} \finer D_{k,l}$ for all $k \leq i$ and for all $j$ and $l$.
+As needed, we may overload the notation by letting $D_{i,j} =  D_{i, j \mod i}$ where $j\mod i$ is the smallest positive residue of $j$ under $i$.
+
+Let $D_{i,j}$ and $D_{k,l}$ be two of these subsets.
+Consider the intersection $D_{i,j}\cap D_{k,l}$ and suppose it is nonempty; let $n$ be in this intersection.
+Then $D_{i,j} = D_{i,n}$ and $D_{k,l} = D_{k,n}$, so the intersection is given by $D_{i,n}\cap D_{j,n} = \{x\in \mathbb{N} \, | \, x \mod \gcd(i,j) = n\} = D_{\gcd(i,j),n}$.
+
+Note also that $D_{i,j}^C = \cup_{k=1,k\neq j}^{i-1}D_{i,k}$.
+
+It follows that any finite combination of unions, intersections, and complements of sets of the form $D_{i,j}$ will be empty, or countably infinite.
+
+Next, let $n\in\mathbb{N}$.
+Consider the union $S_n = \cup_{i=n+1}^{\infty}D_{i,0} = \{n+1,n+2,\ldots\}$, which may be generated by a countable union.
+Thus $S_n^C = \{1,2,\ldots,n\}$.
+By combining complements and intersections, we can obtain $\{n\} = S_{n-1}\cap S_n^C$.
+We can then generate arbitrary subsets of $\mathbb{N}$ using countable operations of union, intersection, and complement on the sets $D_{i,j}$.
+
+Claim: all finite operations of union, intersection, and complement of $D_{i,j}$'s yield sets that may be written as finite unions of $D_{i,j}$'s.
+Countable operations yield arbitrary subsets of $\mathbb{N}$.
+
+Possible source to read from: http://www.stat.cmu.edu/tr/tr814/tr814.pdf
+Note in abstract, gives $1/m$ probability (measure) to $D_{m,j}$ for each $j$.
+
+\section{Insightful failures}
+
+\subsection{Maximal chains}
+
+The idea was to see whether, instead of requiring totality, we could require that we could always find statements at a desired level of description.
+We wanted to say that a maximal chain in narrowness, that is a sequence of statements ordered by narrowness where no other statement could be added, would touch all levels of descriptions, all equivalence classes in equigranularity.
+This does not work.
+
+The issue is that the granularity of a countable sequence ``jumps'' at infinity.
+For example, take the sequence $\{1\}$, $\{1, 3\}$, $\{1, 3, 5\}$, ...
+, \{ all odd natural numbers\}.
+Compare to $\{1\}$, $\{1, 2\}$, $\{1, 2, 3\}$, ...
+, \{ all natural numbers\}.
+They all are maximal chains.
+They both go through all finite sets, but they reach infinity at two different levels.
+Effectively, the above property is only true over finite sets.
 
@@
 
-\begin{prop}\label{uniformconsequence}
-    Let $\tdomain$ be uniform.
-Let $\stmt, \stmt_1 \in \tdomain$ such that $\stmt \sfiner \stmt_1$.
-Then there exist $\stmt_2 \in \tdomain$ such that $\stmt_2 \narrower \stmt_1$ and $\stmt \eqgran \stmt_2$.
-\end{prop}
-
-The start of a proof....
-Let $S = \{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$. We will find a maximal element with respect to fineness (OR narrowness??).
-
-The difficulty here is that for Zorn we need to find an upper bound of an arbitrary chain which also lies in $S$.
-But such an upper bound is entirely unclear on how to obtain without using the proposition to begin with.
-Uniformity is difficult to make much use of for infinite domains (beyond that $S$ is nonempty).
-
-This set $S$ is nonempty, since any possibility narrower than $\stmt_1$ must also be finer than $\stmt$ by uniformity, and the statement $\stmt_1$ must not be the impossibility since it is strictly coarser than the statement $\stmt$.
-
-Now, given any chain (with respect to narrowness), consider its disjunction. But again this may be coarser than $\stmt$!
-
-Note: problem with chains for fineness.
-Taking the disjunction of a chain with respect to fineness can yield something much coarser (i.e.
-outside of $\{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$) if they do not overlap.
-
-Idea: given chains in fineness, can we construct new chains which are, at each level, equigranular, but now they are also chains in narrowness?
-
-Lemma: If $C$ is a chain in fineness and $\stmt$ is such that $\stmt[t]\finer\stmt$ for all $\stmt[t]\in C$, then there exists a set $N$ and a bijection $f:C\to N$ such that $N$ is a chain in narrowness, $\OR N \narrower \stmt$, and for all $\stmt[t]\in C$, $\stmt[t]\eqgran f(\stmt[t])$.
-Use this to move chains in fineness over to chains in narrowness and then can find maximal elements in the set $S$ above ``up to equigranular equivalence classes".
-
-Problem: this is actually much harder. Prop 6 is the case of this Lemma for a chain of length 2 and the top of the chain being equal to $\stmt$.
-
-\section{Notes}
-
-One condition for the measure to be unique given a unit statements is that all points (i.e.
-possibilities in our framework) are equigranular.
-In fact, consider a domain with just three points not equigranular to each other, but each finer than the other.
-We can construct any measure that satisfies the requirement by picking three numbers and assigning them accordingly, since we can't determine ``how bigger'' is one point with respect to the other.
-
-\subsection{Finding a probability measure}
-
-Our first goal is to find a measure which is compatible with the fineness preorder.
-The idea is that fineness tells us which sets are ``larger" (less precise) than each other, regardless of any overlap.
-A measure will quantify this and allow us to assign numbers to the sizes.
-
-The paper \cite{probexistence} seems to be extremely relevant here.
-The basic idea in this paper is to find necessary and sufficient conditions for the existence of a probability measure on a Boolean algebra (e.g.
-of subsets of some set) which is compatible with a total pre-order.
-There are a few points to keep in mind.
-
-First, we need a \emph{total} pre-order.
-This means for every pair of elements, one is finer than the other.
-This is certainly not the case for fineness in general: color and position, for example, cannot be compared.
-This is the goal of defining the \emph{unit} above.
-The unit serves as a baseline statement against which others can be compared for fineness.
-For example, a statement about position singles out length as the dimension to compare, or a statement about numbers of objects specifies ``count" as the dimension to compare (e.g.
-the discrete unit of number of objects).
-
-Next, we need the three properties specified in the paper, to be written up later! We are optimistic that these three properties are present in our situation.
-
-There is one additional caveat.
-The measure constructed by the main theorem in the above-mentioned paper is always a probability measure.
-This means the whole space has finite measure.
-For example, one could put the Lebesgue measure on $(0,1)$ but under a homeomorphism $(0,1)\to\mathbb{R}$ induce a finite (probability) measure on  $\mathbb{R}$.
-If we chose the full set as a unit, we are exactly looking for a probability measure.
-If we pick a unit that is not the full set, we can also "constrain" the problem to that subset, so that we still have a probability measure.
-Then we may be able to "stich together" the measures constructed on all equigranular sets assuming they cover the full space.
-
-\subsection{Applying results of the paper \cite{villegas}}
-
-In this part, we will assume that $\tdomain_{\stmt[u]}$ is a subset of the theoretical domain, $\stmt[u]\in \tdomain_{\stmt[u]}$ is a distinguished element (we'll refer to it as the \emph{unit} when needed), and there is a total pre-order $\leq_{\stmt[u]}$ defined on $\tdomain_{\stmt[u]}$.
-We will write $=_{\stmt[u]}$ for $\leq_{\stmt[u]}$ and $\geq_{\stmt[u]}$, and we write $<_{\stmt[u]}$ for $\leq_{\stmt[u]}$ and not $=_{\stmt[u]}$.
-We will use the word ``equiprobable" for $=_{\stmt[u]}$ since it means something different from equigranular (since $\leq_{\stmt[u]}$ is a different ordering from fineness, even though it will be derived from fineness).
-We assume $\tdomain_{\stmt[u]}$ is closed under complement and countable union (and hence countable intersection), and that there are two additional elements, $\certainty$ and $\impossibility$, which are maximal and minimal in this ordering, respectively.
-
-\begin{remark}
-In defining the subset $\tdomain_{\stmt[u]}$ and the total preorder $\leq_{\stmt[u]}$, we will have to be very careful.
-We want to ``squash" together all the sets which should be measure zero in this unit (e.g.
-anything ``lower-dimensional"), ``blow up" all the ``infinite measure" subsets (e.g.
-``higher dimensional"), and ensure that everything is compatible under this order, so there can be only one ``type" of property under consideration.
-\end{remark}
-
-We write out the main axioms and statements from \cite{villegas} which interest us in this situation, in the notation established earlier. On page 1789, we have the following:
-
-Axiom of Pre-order:
-\begin{itemize}
-    \item Totality: If $\stmt_1,\stmt_2\in \tdomain_{\stmt[u]}$, then $\stmt_1 \leq_{\stmt[u]} \stmt_2$ or $\stmt_2\leq_{\stmt[u]} \stmt_1$
-    \item Reflexivity: For all statements $\stmt\in \tdomain_{\stmt[u]}$, we have $\stmt\leq_{\stmt[u]}\stmt$
-    \item Transitivity: For all statements $\stmt_1,\stmt_2,\stmt_3\in \tdomain_{\stmt[u]}$, if $\stmt_1\leq_{\stmt[u]}\stmt_2$ and $\stmt_2\leq_{\stmt[u]}\stmt_3$, then $\stmt_1\leq_{\stmt[u]}\stmt_3$.
-    \item Existence of max and min: We have $\impossibility\leq_{\stmt[u]}\certainty$ and \emph{not} $\certainty\leq_{\stmt[u]}\impossibility$.
-For all $\stmt\in \tdomain_{\stmt[u]}$ we have $\impossibility\leq_{\stmt[u]}\stmt\leq_{\stmt[u]}\certainty$
-\end{itemize}
-
-The above list is simply the requirement that $\leq_{\stmt[u]}$ is a total pre-order and that there are a maximal and minimal element.
-
-Axiom of Monotony: 
-Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
-Assume that $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, and $\stmt_2\ncomp \stmt_4$ (i.e.
-$\stmt_2\AND \stmt_4 =_{\stmt[u]} \impossibility$).
-Then we have $\stmt_1\OR \stmt_3 \leq_{\stmt[u]} \stmt_2\OR \stmt_4.$
-
-Monotony guarantees that given two incompatible statements, each ``bigger" than another statement, then their disjunction is still bigger than the disjunction of the smaller statements.
-
-The following are consequences of the two axioms above. Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
-\begin{itemize}
-    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ and $\stmt_2<_{\stmt[u]} \stmt_3$, then $\stmt_1<_{\stmt[u]}\stmt_3$. 
-    \item If $\stmt_1\narrower \stmt_2$ then $\stmt_1\leq_{\stmt[u]} \stmt_2$
-    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, $\stmt_2\leq_{\stmt[u]} \stmt_4$, and $v\leq_{\stmt[u]} \stmt_1$, then $\stmt_2 \AND \NOT \stmt_1 \leq_{\stmt[u]} \stmt_4 \AND \NOT \stmt_3$
-    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ then $\NOT \stmt_2 \leq_{\stmt[u]}\NOT \stmt_1$
-\end{itemize}
-
-Next, we will need a more subtle axiom in order to make sense of infinite combinations of statements.
-
-\begin{remark}
-The \emph{fineness} relation $\finer$ should satisfy all of the axioms of pre-order and the axiom of monotony, except for totality. We restrict to $\tdomain_{\stmt[u]}$ in order to get totality. 
-\end{remark}
-
-Axiom of Monotone Continuity:
-Let $\{\stmt_i : i\in \mathbb{Z}^+\}$ be a sequence of statements which is monotone increasing with respect to narrowness (containment), and let $\stmt = \cup_i\stmt_i$ be the disjunction.
-Let $t\in \tdomain_{\stmt[u]}$ such that $\stmt_i\leq_{\stmt[u]} t$ for all $i$.
-Then $\stmt\leq_{\stmt[u]} t$.
-
-In other words, an infinite disjunction can only strictly exceed a statement in coarseness if the disjunction of a finite subset of them already exceeds it.
-Here is an interesting consequence of the combination of the above axioms:
-
-\begin{prop}[Corollary to Lemma 1 in \cite{villegas}]
-\label{infinitenull}
-Let $\{\stmt_i : i\in \mathbb{Z}^+\}$ be an infinite collection such that $\stmt_i\AND\stmt_j =_{\stmt[u]} \impossibility$ for all $i\neq j$, and $\stmt_i =_{\stmt[u]} \stmt_j$ for all $i,j$.
-Then it must be the case that $\stmt_i=_{\stmt[u]}\impossibility$ for all $i$.
-\end{prop}
-This says that infinite collections of equiprobable statements with null overlaps must be null themselves.
-
-An \emph{atom} is a statement $\stmt\in S$ such that $\impossibility<_{\stmt[u]}\stmt$ and there does not exist any statement $\stmt_2\in S$ such that $\impossibility<_{\stmt[u]} \stmt_2<_{\stmt[u]} \stmt$.
-In particular, Theorem 5 in \cite{villegas} \S3 includes the fact that $\tdomain_{\stmt[u]}$ is atomless if and only if every statement can be partitioned into two equiprobable statements.
-Theorem 3 in \S4 states that if $\tdomain_{\stmt[u]}$ is atomless, then there exists a unique probability measure compatible with $\leq_{\stmt[u]}$, and it is countably additive!
-
-\subsection{Using Monotone Continuity}
-
-The core of the proof of Theorem 3 in Section 4 of \cite{villegas} comes from Theorem 5 (iv) in Section 3.
-In the terminology of Villegas, a \emph{qualitative probability $\sigma$-algebra} is a $\sigma$-algebra of subsets of a space with an ordering that satisfies the Axiom of Pre-order, the Axiom of Monotony, and the Axiom of Monotone Continuity.
-We also need the language of \emph{atoms}, defined earlier.
-We will reproduce some of the ideas in the proofs of these theorems in the language of statements and our theoretical domain, while attempting to keep track of the usages of Monotone Continuity.
-The hope is that there is a weakened version of the axiom that gets us the desired result, or that we can build the ordering $\leq_{\stmt[u]}$ with Monotone Continuity but it can still distinguish the infinite sets described earlier.
-
-Let $A$ be a statement in $\tdomain_{\stmt[u]}$ with $A>_{\stmt[u]} \impossibility$. A \emph{minor incomplete partition} of $A$ is a collection $M_A$ of statements with the following properties:
-\begin{itemize}
-    \item We have $\stmt>_{\stmt[u]}\impossibility$ and $\stmt \narrower A$ for all $\stmt\in M_A$, that is, they are non-null and narrower than $A$
-    \item For all $\stmt_1,\stmt_2\in M_A$, we have $\stmt_1\AND\stmt_2 =_{\stmt[u]} \impossibility$, that is, they overlap only on statements ``squashed" to zero. 
-    \item The disjunction of $M_A$ is finer than its complement in $A$: $$\bigOR_{\stmt\in M_A}\stmt \finer (A\AND \NOT\bigOR_{\stmt\in M_A}\stmt).$$ This is the condition of being \emph{minor}.
-    %UM_A finer than A - UM_A^C
-\end{itemize}
-
-It is shown in \cite{villegas} that any (minor) incomplete partition is at most countable (Theorem 3).
-Rather than give the full proof of this fact, we note that it follows from a large collection of lemmas, the most interesting of which is the following, which we will prove below:
-
-\begin{prop}[Lemma 1 from \cite{villegas}]
-If $\{\stmt[a]_i: i\in I\}$ is an infinite set of statements such that (1) $\stmt[a]_i \AND \stmt[a]_j =_{\stmt[u]}\impossibility$ for all $i\neq j$ and (2) $\stmt[a]_i>_{\stmt[u]}\impossibility$ for all $i$, then the following holds.
-Given a statement $\stmt>_{\stmt[u]}\impossibility$, only finitely many of the $\stmt[a]_i$'s have $\stmt[a]_i>_{\stmt[u]}\stmt$.
-\end{prop}
-\begin{proof}
-By way of contradiction, let $\stmt[a]_1,\ldots$ be a sequence of statements violating the conclusion (i.e.
-infinitely many statements $\stmt[a]_i$ with $\stmt[a]_i>_{\stmt[u]}\stmt>_{\stmt[u]}\impossibility$.
-Consider the new sequence $$\stmt_i = \bigOR\limits_{n=i}^{\infty}\stmt[a]_n.$$ This sequence is monotonely decreasing to the impossibility in narrowness (and hence in fineness and in $>_{\stmt[u]}$), but we also have $\stmt_i\geq_{\stmt[u]}\stmt$ for all $i$.
-By monotone continuity (the contrapositive), we have $\stmt[s]\leq_{\stmt[u]}\impossibility$, a contradiction.
-\end{proof}
-
-A quick corollary of this is Proposition \ref{infinitenull}, above. Using minor incomplete partitions, we prove the following.
-
-\begin{prop}[Theorem 4, \S3 \cite{villegas}]
-\label{split}
-If there are no atoms, then for every statement $\stmt$ there exist statements $\stmt_1$ and $\stmt_2$ such that: $\stmt_1\ncomp \stmt_2$, $\stmt_1\equ\stmt_2$, and $\stmt_1\OR\stmt_2 = \stmt$.
-In other words, $\stmt$ can be partitioned into two equiprobable (note: I remember we don't like this word but I don't remember if we replaced it or not) statements.
-\end{prop}
-\begin{proof}
-Let $S$ be the set of minor incomplete partitions of $\stmt$, and endow it with a partial order by inclusion.
-Given a totally ordered chain in $S$, its union is still a minor incomplete partition, and is thus an upper bound for the chain.
-We may apply Zorn's lemma to find a maximal element of $S$, the disjunction of which we call $\stmt_1$.
-We claim that $\stmt_1$ and $\stmt_2:= \stmt\AND\NOT\stmt_1$ are equiprobable, incompatible statements.
-Incompatibility is immediate from the definition.
-For equiprobability, first note that by definition of being a minor incomplete partition we have $\stmt_1\lequ\stmt_2$.
-Next, if $\stmt[t]_1,\ldots$ is a monotone decreasing (with respect to $\lequ$) sequence of statements $\stmt[t]_n>_{\stmt[u]}\impossibility$ narrower than $\stmt_2$ and converging to a zero-comparable event, we have that $\stmt_1\OR\stmt[t]_n>_{\stmt[u]}\stmt_2\OR\stmt[t]_n$ since $\stmt_1$ is maximal.
-We needed no atoms in order to guarantee this sequence exists (Lemma 3 in Section 3 of \cite{villegas}).
-It follows by monotone continuity that $\stmt_1\geq_{\stmt[u]} \stmt_2$.
-\end{proof}
-
-We will also need a few types of \emph{random variables} as defined in \cite{villegas} \S3.
-Let $A({\stmt[u]}) = \{x \in X : x \comp \stmt[u]\}$ be the set of possibilities associated with $\stmt[u]$.
-This is a subset of the collection of possibilities which will serve as the ``sample space" after picking a unit $\stmt[u]$.
-
-A \emph{uniform sequence of finitely-valued random variables} is a sequence of functions (random variables in the traditional sense) $X_1,X_2,\ldots: A({\stmt[u]})\to E$ where $E$ is a finite set, with the following property: for all $\{x_1,\ldots,x_n\}\subseteq E$, all the events $$\{p\in A({\stmt[u]}): X_i(p) = x_i, i=1,\ldots,n\}$$
-are equiprobable ($\equ$ with each other).
-
-Notice that monotone continuity allows us to let $n\to\infty$ to define events where the whole sequence is specified with the resulting events all being equiprobable.
-By Proposition \ref{infinitenull} above, all such events are null.
-
-Building on these sequences of random variables, we will use \emph{uniform random variables} to define probability measures.
-Let $X:A({\stmt[u]})\to[0,1]$ be a random variable, and let $I$ and $J$ be intervals contained in $[0,1]$.
-We say $X$ is a uniform random variable if $X^{-1}(I)\lequ X^{-1}(J)$ if and only if $|I|\leq|J|$ for all such $I,J$.
-
-From Proposition \ref{infinitenull}, $X^{-1}(a)$ is a zero-comparable event for all $a\in[0,1]$, and if $0=a_0<a_1<\cdots<a_n=1$, then $X^{-1}([a_i,a_{i+1}))$ are a finite partition of $\tdomain$.
-
-Next, we will pull many of the above concepts together to show how we obtain a probability measure on $A({\stmt[u]})$ realized as a function on $\tdomain$ compatible with $\lequ$.
-
-\begin{prop}\label{uniformexist}
-If $\tdomain_{\stmt[u]}$ has no atoms with respect to $\lequ$, then there exists a uniform random variable.
-\end{prop}
-\begin{proof}
-By Proposition \ref{split} we may split any statement into two equiprobable, null-compatible sets.
-Starting with the statement $\stmt[u]$ (here, the ``universe" which we want to have probability 1), repeatedly divide into two statements in stages.
-At stage $n$ (starting with $n=0$), we will have statements $\stmt[u]^{(n)}_1,\ldots,\stmt[u]^{(n)}_{2^n}$.
-Now, take the disjunction of alternating statements in each collection so that there is a ``right" and ``left" (odd and even index) half at each stage.
-For example, at stage $n=1$, the odd and even halves are simply $\stmt[u]^{(1)}_1$ and $\stmt[u]^{(1)}_2$; at stage $n=2$ we have $\stmt[u]^{(2)}_1\OR\stmt[u]^{(2)}_3$ and $\stmt[u]^{(2)}_2\OR\stmt[u]^{(2)}_4$.
-
-Define a sequence of random variables $X_n:A(\stmt[u])\to\{0,1\}$ by $X_n(a) = 1$ if and only if $a$ is in the odd half in stage $n$.
-This gives each $a\in A(\stmt[u])$ a binary ``address".
-By the construction, it follows that this is a uniform sequence of finitely-valued random variables (defined above).
-
-Now, we define a uniform random variable $X:A(\stmt[u])\to[0,1]$.
-For $a\in A(\stmt[u])$, define $X(a) = \sum_{n=1}^{\infty}2^{-n}X_n(a)$.
-In other words, $X(a)$ is the binary expansion $0.X_1(a)X_2(a)\ldots_2$.
-
-One can show immediately that $X^{-1}(I)\lequ X^{-1}(J)$ if and only if $|I|\leq |J|$ when $I$ and $J$ have endpoints on numbers with finite binary expansions (since $X_1,\ldots,X_n$ are uniform), and by monotone continuity, the uniformity condition extends to arbitrary intervals in $[0,1]$.
-\end{proof}
-
-Finally, we reach our goal of building a probability measure on $A(\stmt[u])$.
-
-\begin{prop}
-If $\tdomain_{\stmt[u]}$ has no atoms with respect to $\lequ$, then there exists a unique compatible probability measure, and it is countably additive. 
-\end{prop}
-\begin{proof}
-By Propsition \ref{uniformexist}, let $X:POSS_{\stmt[u]}\to[0,1]$ be a uniform random variable.
-%Define $F:[0,1]\to\tdomain$ by $F(x) = X^{-1}([0,x])$.
-It is straightforward to see that $F$ is monotone in $x$ (with respect to $\lequ$), and that if $x_n\to x$, then $\lim_nF(x_n) \equ F(x)$.
-%Started writing this part from paper, but it is not directly used in construction.
-Will add details later as needed.
-We claim that for any statement $\stmt\in\tdomain_{\stmt[u]}$, there exists a unique number $p\in[0,1]$ such that $\stmt\equ X^{-1}([0,p])$.
-Define $P:\tdomain_{\stmt[u]}\to[0,1]$ by $P(\stmt)= p$.
-This defines a probability measure compatible with $\lequ$.
-Details forthcoming as needed....
-\end{proof}
-
-Ideas involved: Defines a function $X:A(\stmt[u])\to [0,1]$ in a way that is consistent with $\lequ$ by having ``larger" sets of possibilities (i.e.
-coarser statements) corresponding to larger intervals in $[0,1]$.
-Then, the measure of any other statement is given by finding an equiprobable statement which is the preimage of $[0,p]$ under $X$ and defining the measure to be $p$.
-
-Wish: find a way to prove the above statements in an order-theoretic way.
-In particular, we wish to show that the order type of the equigranular-equivalence classes is equivalent to (a subset of) $\mathbb{R}$.
-Sufficient conditions to be isomorphic to $\mathbb{R}$ are (1) lacking endpoints (2) complete (i.e.
-has Least Upper Bound property) (3) has dense countable subset.
-The use of random variables may be very similar to an approach using order theory.
-For example, the transition from a uniform sequence of finitely-valued random variables into a (single) uniform random variable makes use of a countable dense subset.
-
-\subsection{Example with lack of monotonicity}
-
-Consider a Euclidian two dimensional space charted by $\{x,y\}$.
-Let $U_{i,j}$ be the square in the interval $x \in (i, i+1)$ and $y \in (j, j+1)$.
-Let $V$ be the vertical (infinite) strip in the interval $x \in (0, 1)$ and $W$ the vertical strip in the interval $x \in (0, 2)$.
-Define the sequence $W_i = \{ U_{0,0}, U_{1,0}, U_{0,1}, U_{1,1}, U_{0,2}, U_{1,2}, U_{0,3}, U_{1,3}, ...
-\}$.
-We have $\bigcup\limits_0^\infty W_i = W$.
-This means $\bigcup\limits_0^n W\finer V$ for all finite $n$ but $\bigcup\limits_0^\infty W_i = W \coarser V$.
-
-\subsection{Some thoughts on Example in 3.3}
-
-The construction above hints at some fundamental concepts in measure theory which will be important for us to consider.
-
-Notice that the two sets $V$ and $W$ are both of infinite Lebesgue measure (area).
-With respect to Lebesgue measure, monotone continuity is immediate since if $\bigcup\limits_0^n W\finer V$ for all $n$ then $\lim\limits_{n\to\infty}\mu(\bigcup\limits_0^n W) = \mu(W)\finer V$.
-Even the above example does not violate this, since there, both would be ``equivalent" in fineness, namely they are infinitely coarse.
-
-Our goal is to have a measure which is able to distinguish sets which may be of infinite Lebesgue measure as well.
-This will be very difficult to fully capture.
-Consider the following procedure.
-Starting with the vertical strip $V$, slide all the squares at even $y$-values to the right by one, and then allow the squares to collapse down to rest on the square below.
-Now, by simply rearranging regions, we have (almost) a cover of the wider strip $W$.
-The only missing points are the line $x=1, y>0$.
-If instead we used $(0,1]$ and $(0,2]$ for the strips, this procedure would cover them exactly.
-One could continue and cover the entire plane with similar translations.
-
-Here is one approach to distinguish $V$ and $W$, and similar methods will enable one to distinguish many more infinite-measure subsets.
-Let $S_n$ be the $2n\times 2n$ square centered the origin with vertices at $(\pm n, \pm n)$.
-Then for all $n$, $V\cap S_n$ has area $n$, and $W\cap S_n$ has area $2n$.
-This holds for all $n$ and so we might use something like this to think about how we would distinguish infinite areas.
-Notice that this is in some sense writing $V$ and $W$ as countable unions of nested sets in a standardized way (e.g.
-$V = \bigcup\limits_n V\cap S_n$) and then comparing the finite-area subsets.
-One could choose shapes other than $S_n$ for this purpose.
-Another way to distinguish them would be to look at the limit $\lim\limits_{n\to\infty}\text{Area}(V\cap S_n)/\text{Area}(S_n)$; in this case they are both 0 but for a different choice of $S_n$ one could find an asymptotic difference.
-The point of the above discussion is to try to gain a better idea of which (infinite/unbounded) sets we want to be related by $\finer$.
-It is not clear even for ``nice" regions of $\mathbb{R}^2$, let alone at the full generality we are seeking.
-
-Regardless, the goal is not just to distinguish infinite-Lebesgue-measure sets. We need to understand when monotone continuity is relevant and what happens without it.
-
-\subsection{Infinitesimals}
-
-\subsubsection{Notes on deFinetti and Archimedean property }
-
-On Page 21 de Finetti specifically says that $P(E) \geq P(E’)$ implies $E \geq E’$ but not vice-versa.
-And that we can have $P(A)=P(B)=0$ even if $A<B$ if $A$ is impossible and $B$ is not.
-So, the ordering he has in mind is fineness for us.
-And then he says that this means that the 4 postulates give a non-Archimedean structure.
-But you can make it Archimedean by ignoring the probabilities that are “infinitesimally small”: those that multiplied by a number, however big, will never become the certainty.
-This whole thing is spot on.
-
-From a cursory look at the Archimedean property (\url{https://en.wikipedia.org/wiki/Archimedean_property})
-this is typically discussed to define notions of infinitesimals in an ordered group. Most notably, the real number (and any complete ordering) will satisfy the Archimedean property.
-
-\subsubsection{Infinitesimals}
-
-Having seen that, we can define infinitesimal in the following way.
-
-\begin{defn}\label{infinitesimals}
-Let $\stmt[u],\stmt[s] \in \tdomain$.
-Then $\stmt$ is \textbf{infinitesimal} with respect to $\stmt[u]$ if for any natural number $n$ we can find a set of $n$ statements $\{ \stmt_i \}_{i=1}^n$ such that:
-\begin{enumerate}
-    \item $\stmt \finer \stmt_i$ for all $i$
-    \item $\stmt_i \ncomp \stmt_j$ for all $i \neq j$
-    \item $\bigOR\limits_{i=1}^n \stmt_i \finer \stmt[u]$ 
-\end{enumerate}
-In words, $\stmt$ is infinitesimal with respect to $\stmt[u]$ if for any number, there is a disjunction of that many broader mutually incompatible statements $\stmt_i$ is narrower than $\stmt[u]$.
-\end{defn}
-
-\begin{coro}
-    The impossibility $\impossibility$ is an infinitesimal with respect to any $\stmt[u] \in \tdomain$.
-\end{coro}
-
-\begin{proof}
-    Let $\stmt[u] \in \tdomain$ and consider $\{ \stmt_i \}_{i=1}^n$ where $\stmt_i \equiv \impossibility$ for all $i$.
-Then $\impossibility \finer \impossibility \equiv \stmt_i$ for all $i$.
-Since $\impossibility \ncomp \impossibility$, $\stmt_i \ncomp \stmt_j$ for all $i$ and $j$.
-Lastly, $\bigOR\limits_{i=1}^n \stmt_i \equiv \impossibility \finer \stmt[u]$.
-This means that $\{ \stmt_i \}_{i=1}^n$ realizes infinitesimality for $\impossibility$, which is an infinitesimal of $\stmt[u]$.
-\end{proof}
-
-\begin{prop}\label{infinitesimal_conjunction}
-Let $\stmt[v] \in \tdomain$ be an infinitesimal with respect to $\stmt[u] \in \tdomain$. Then any $\stmt \finer \stmt[u]$ is also an infinitesimal with respect to $\stmt[u]$.
-\end{prop}
-\begin{proof}
- Since $\stmt[v]$ is an infinitesimal with respect to $\stmt[u]$, for any $n$ we can find a set of statements $\{ \stmt_i \}_{i=1}^n$ that realizes its infinitesimality.
-We have $\stmt \finer \stmt[v] \finer \stmt_i$ for all $i$.
-The other two properties do not depend on $\stmt[v]$.
-Therefore $\{ \stmt_i \}_{i=1}^n$ realizes infinitesimality for $\stmt$ as well so $\stmt$ is an infinitesimal with respect to $\stmt[u]$.
-\end{proof}
-
-\begin{coro}
-The conjunction of two infinitesimals with respect to $\stmt[u]$ is infinitesimal with respect to $\stmt[u]$.
-\end{coro}
-\begin{proof}
-Let $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ be infinitesimal with respect to $\stmt[u]$.
-Since $\stmt[v]_1 \AND \stmt[v]_2 \narrower \stmt[v]_1$, $\stmt[v]_1 \AND \stmt[v]_2 \finer \stmt[v]_1$ and therefore $\stmt[v]_1 \AND \stmt[v]_2$ by \ref{infinitesimal_conjunction}.
-\end{proof}
-
-\begin{prop}
-The disjunction of two infinitesimals with respect to $\stmt[u]$ that are comparable\footnote{Is comparability required?
-Can it be proved one way or another?} to each other is infinitesimal with respect to $\stmt[u]$.
-\end{prop}
-\begin{proof}
-Let $\stmt[v],\stmt[w]$ be infinitesimal with respect to $\stmt[u]$.
-Without loss of generality, suppose $\stmt[v] \finer \stmt[w]$.
-Let $n>0$.
-Let $\{\stmt[w]_i\}_{i=1}^{2n}$ be a collection realizing infinitesimality for $2n$ for $\stmt[w]$.
-Define $\stmt_j = \stmt[w]_j\OR\stmt[w]_{j+n}$ for $j=1,\ldots,n$.
-Notice that the $\stmt_j$'s are mutually incompatible, and their disjunction must be finer than $\stmt[u]$.
-Further, since $\stmt[v]\finer\stmt[w]$ we have that $\stmt[v]\finer\stmt[w]_i$ for all $i$.
-It follows that:
-$$
-\stmt[v]\OR\stmt[w] \finer \stmt[w]_j\OR\stmt[w]_{j+n} = \stmt_j
-$$
-for all $j=1,\ldots,n$.
-Thus the collection $\{\stmt_j\}_{j=1}^n$ realizes $\stmt[v]\OR\stmt[w]$ being infinitesimal with respect to $\stmt[u]$ for $n$.
-We conclude that $\stmt[v]\OR\stmt[w]$ is infinitesimal with respect to $\stmt[u]$.
-\end{proof}
-
-\begin{prop}[Infinitesimality is consistent with fineness]
-Let $\stmt[u],\stmt[v]$ be statements such that $\stmt[v]\finer\stmt[u]$.
-Let $\stmt$ be a statement $\stmt\narrower\stmt[u]\AND\stmt[v]$ such that $\stmt$ is infinitesimal with respect to $\stmt[v]$.
-Then $\stmt$ is infinitesimal with respect to $\stmt[u]$.
-\end{prop}
-\begin{proof}
-Let $n>0$ be a natural number, and let $\{ \stmt_i \}_{i=1}^n$ be as in the definition of infinitesimal for $\stmt$ with respect to $\stmt[v]$.
-Then $\{ \stmt_i \}_{i=1}^n$ also satisfies the requirements for infinitesimal with respect to $\stmt[u]$.
-The first two requirements are unchanged, and the third is easy to see:
-$$
-\bigOR\limits_{i=1}^n \stmt_i \finer\stmt[v]\finer \stmt[u]
-$$
-\end{proof}
-
-As an immediate corollary to the above, we have:
-\begin{prop}
-Infinitesimality is consistent with $\narrower$.
-\end{prop}
-
-We wish to also show some form of compatibility in the reverse direction.
-That is, if we shrink by a ``finite" amount, the sets which were previously infinitesimal should remain so (of course, if we shrink by an ``infinite" amount, e.g.
-by dropping a dimension, previously infinitesimal sets may no longer be so).
-
-\begin{prop}
-Let $\stmt[u],\stmt[v]$ be statements such that $\stmt[v]$ is not infinitesimal with respect to $\stmt[u]$.
-Let $\stmt$ be a statement $\stmt\narrower\stmt[v]$ such that $\stmt$ is infinitesimal with respect to $\stmt[u]$.
-Then $\stmt$ is infinitesimal with respect to $\stmt[v]$.
-\end{prop}
-\begin{proof}
-Let $\stmt[u]$, $\stmt[v]$, and $\stmt$ be as in the Proposition statement.
-We will show that $\stmt$ is infinitesimal with respect to $\stmt[v]$.
-To do so, we will show that for any $n>0$, we can find a set of $n$ statements $\{\stmt_i\}_{i=1}^n$ such that:
-\begin{enumerate}
-    \item $\stmt \finer \stmt_i$ for all $i$
-    \item $\stmt_i \ncomp \stmt_j$ for all $i \neq j$
-    \item $\bigOR\limits_{i=1}^n \stmt_i \finer \stmt[v]$ 
-\end{enumerate}
-We will show this for arbitrary $n$, which will be sufficient to prove that $\stmt$ is infinitesimal with respect to $\stmt[v]$.
-
-First, since $\stmt[v]$ is not infinitesimal with respect to $\stmt[u]$, there exists some $N>0$ such that for any set of $N$ statements, at least one of the conditions of infinitesimality of $\stmt[v]$ with respect to $\stmt[u]$ fails.
-
-Let $n>0$.
-Let $\{\stmt_i\}_{i=1}^{nN}$ be a set of $n\cdot N$ statements realizing infinitesimality of $\stmt$ with respect to $\stmt[u]$ (i.e.
-they are mutually incompatible, broader than $\stmt$, and their disjunction is narrower than $\stmt[u]$).
-
-{\bf By totality of $\finer$}\footnote{What we needed here is that there is a partition of the collection into $N$ sets of size $n$ such that for one of them, their disjunction is finer than each of $N-1$ disjunctions of the other subsets in the partition.}, we may assume the collection $\{\stmt_i\}_{i=1}^{nN}$ is ordered by fineness ($\stmt_1\finer\stmt_2\finer\cdots$).
-Consider the finest $n$ statements, that is, consider the collection $\{\stmt_i\}_{i=1}^{n}$.
-We will show that this collection satisfies the conditions for $\stmt$ being infinitesimal with respect to $\stmt[v]$ for $n$ statements.
-
-By construction, this satisfies properties 1 and 2 for infinitesimality of $\stmt$ in $\stmt[v]$, since it is a subset of a collection which already satisfied those properties.
-
-Now suppose it fails property 3\footnote{We will use totality since we need $\NOT \big(\stmt[x]\sfiner\stmt[y]\big) \implies \stmt[x]\coarser\stmt[y]$.}, that is, suppose 
-$$
-\bigOR\limits_{i=1}^n \stmt_i \scoarser\stmt[v].
-$$
-We will show that this will lead to a contradiction by implying that there is a collection of statements of size $N$ realizing $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$.
-
-Consider now the collection of $N$ statements $\{\stmt[a]_j\}_{j=1}^N$ defined by $$\stmt[a]_j = \bigOR\limits_{i=n(j-1)+1}^{nj}\stmt_i.$$
-These statements are built from the collection $\{\stmt_i\}_{i=1}^{n}$ by taking $N$ disjunctions each consisting of $n$ of the statements.
-Notice that $\stmt[a]_1$ is the disjunction of the collection $\{\stmt_i\}_{i=1}^{n}$ considered above.
-
-First, this collection is mutually incompatible, since it is made up of disjunctions of mutually incompatible statements.
-This is property 2 for $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$.
-
-Next, these must satisfy property 1 for $\stmt[v]$ being infinitesimal with respect to $\stmt[u]$.
-To see why, notice that by our assumption, we have $\stmt[a]_1\scoarser\stmt[v]$, and by construction, we have $\stmt[a]_j\coarser\stmt[a]_1$ for all $j=1,\ldots,N$, and so $\stmt[v]\finer\stmt[a]_j$ for all $j=1,\ldots,N$.
-
-Finally, they also must satisfy property 3: $$\stmt[a]_1\OR\cdots\OR\stmt[a]_N = \bigOR\limits_{i=1}^{nN}\stmt_i\finer\stmt[u].$$
-
-Thus, this new collection violates the assumption that no set of $N$ statements satisfies all three properties for $\stmt[v]$ being infinitesimal in $\stmt[u]$.
-
-{\bf By totality of $\finer$},\footnote{Here we are using totality as specified in the previous footnote.} we conclude that we must have $\bigOR\limits_{i=1}^n \stmt_i \finer\stmt[v]$ as required, and so the collection $\{\stmt_i\}_{i=1}^{n}$ realizes $\stmt$ being infinitesimal with respect to $\stmt[v]$ for $n$ statements.
-Since $n$ was arbitrary, it must be the case that $\stmt$ is infinitesimal with respect to $\stmt[v]$.
-\end{proof}
-
-\subsubsection{Equivalence at u scale}
-
-\def\equivu{\cong_{\stmt[u]}}
-
-\begin{defn}
-Let $\stmt_1, \stmt_2 \in \tdomain$.
-We say $\stmt_1$ is \textbf{equivalent to $\stmt_2$ at $\stmt[u]$ scale}, noted $\stmt_1 \equivu \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$.
-\end{defn}
-
-\begin{prop}
-Equivalence at $\stmt[u]$ scale is an equivalence class.
-\end{prop}
-\begin{proof}
-For reflexivity, let $\stmt \in \tdomain$.
-We have $\stmt \OR \impossibility \equiv \stmt \OR \impossibility$.
-Since $\impossibility$ is an infinitesimal for all $\stmt[u]$ (TODO: add proposition), $\stmt \equivu \stmt$.
-For symmetry, note that the definition itself is symmetrical.
-For transitivity, let $\stmt_1 \equivu \stmt_2 \equivu \stmt_3$.
-Then we have $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$ and $\stmt_2 \OR \stmt[w]_2 \equiv \stmt_3 \OR \stmt[w]_3$.
-\end{proof}
-
-\subsubsection{Greatness in $\stmt[u]$}
-
-First define ``as big in $\stmt[u]$'' $\equ$.
-
-\begin{defn}
-Let $\stmt_1, \stmt_2 \in \tdomain$.
-Then $\stmt_1$ is \textbf{as big as} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \equ \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$.
-\end{defn}
-
-\begin{prop}
-Suppose $\tdomain$ is uniform.
-Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$, then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
-\end{prop}
-
-We start with a simpler case:
-
-\begin{prop}
-Let $\stmt_1,\stmt,\stmt[v]_1\in\tdomain$ such that $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$, and suppose $\stmt[v]_1$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt\equ\stmt_1$. 
-\end{prop}
-\begin{proof}[Using Prop \ref{uniformconsequence}]
-If both of the $\finer$'s are actually equigranular, there is nothing to show.
-Next consider the case that $\stmt_1\sfiner\stmt\finer\stmt_1\OR\stmt[v]_1$.
-Then by Proposition \ref{uniformconsequence} there exists $\stmt_2$ such that $\stmt_2\narrower\stmt$ and $\stmt_2\eqgran\stmt_1$.
-But how do we put this together??
-What about the even simpler version of the lemma with $\narrower$ instead of $\finer$?
-Just use $\stmt\OR (\stmt_1\OR\stmt[v]_1 \AND \NOT\stmt)$.
-\end{proof}
-
-This seems problematic already:
-
-We have $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$.
-We need to find some $\stmt[w],\stmt[w]_1$ infinitesimal with respect to $\stmt[u]$, such that $\stmt\OR\stmt[w]\eqgran\stmt_1\OR\stmt[w]_1$.
-Since we don't know anything about the overlap of $\stmt$ and $\stmt_1$, we can't use any operations like $\stmt\AND\NOT\stmt_1$ to find something small.
-We need to know that additional statements exist in order to find $\stmt[w]$ and $\stmt[w]_1$.
-
-For example, what if every infinitesimal statement was narrower $(\narrower)$ than $\stmt$?
-In this case, we could have something like $\stmt_1\sfiner\stmt\sfiner\stmt_1\OR\stmt[x]$ for all statements (including all infinitesimals) $\stmt[x]$.
-It would then not be the case that $\stmt\equ\stmt_1$.
-
-\begin{prop}[Need something like this but not clear if correct!]
-Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$. Then there exist statements  $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$, infinitesimal in $\stmt[u]$, such that:
-$$
-\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
-$$
-\end{prop}
-
-Note that the existence of the right combination of incompatible statements is sufficient.
-However, that is stronger than we need and may be harder to prove.
-Problem: we have very few propositions about existence of statements!
-It is hard to ensure certain ones exist.
-
-Consider the following (extreme) case. Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$, but the infinitesimals which realize these relationships are $\stmt[v]_1$ and $\stmt[v]_2$, as follows:
-$$
-\stmt_1\OR\stmt[v]_1 \eqgran \stmt_2\OR\stmt[v]_2,\ \stmt_2\OR\stmt[v]_1\eqgran\stmt_3\OR\stmt[v]_2
-$$
-Without further axioms, it may not be possible to put this together to get transitivity (i.e. to find that $\stmt_1\equ\stmt_3$).
-
-\begin{prop}
-The relation $\equ$ is an equivalence relation.
-\end{prop}
-\begin{proof}
-Reflexivity and symmetry are immediate since $\eqgran$ is an equivalence relation.
-For transitivity, suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$.
-Let $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$ be infinitesimals with respect to $\stmt[u]$, such that $\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2$ and $\stmt_2\OR\stmt[w]_2\eqgran\stmt_3\OR\stmt[w]_3$.
-Then we have:
-$$
-\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
-$$
-Since $\stmt[v]_1\OR\stmt[w]_2$ and $\stmt[w]_3\OR\stmt[v]_2$ are infinitesimal with respect to $\stmt[u]$ (being finite disjuncts of infinitesimals), we have $\stmt_1\equ\stmt_3$. 
-\end{proof}
-
-\begin{defn}\label{smalleru}
-Let $\stmt_1, \stmt_2 \in \tdomain$.
-Then $\stmt_1$ is \textbf{smaller than} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \lequ \stmt_2$, if $\stmt_1 \finer \stmt_2$ or $\stmt_1 \equ \stmt_2$.
-We define $\stmt_1\leu\stmt_2$ by $\stmt_1\lequ\stmt_2$ and not $\stmt_1\equ\stmt_2$.
-\end{defn}
-{\bf Warning:} Transitivity doesn't seem to work as stated (see the proof attempt below).
-We propose the following modification: say $\stmt_1\lequ\stmt_2$ if there exist statements $\stmt_1'$ and $\stmt_2'$ with $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ such that $\stmt_1'\finer\stmt_2'$.
-Note that {\bf by totality}, it is sufficient to just say $\stmt_1\finer\stmt_2'$ (no need for $\stmt_1'$).
-
- \begin{prop}
-$\stmt_1 \leu \stmt_2$ if for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$. 
-\end{prop}
-\begin{proof}
-Since $\impossibility$ is infinitesimal, we have $\stmt_1 = \stmt_1\OR\impossibility \sfiner \stmt_2$, and so by Definition \ref{smalleru}, we have that $\stmt_1\lequ\stmt_2$.
-Now, we must show it is not the case that $\stmt_1\equ \stmt_2$.
-
-Let $\stmt[v]_1,\stmt[v]_2\in\tdomain$ be infinitesimals.
-Then $\stmt_1\OR\stmt[v]_1 \sfiner \stmt_2 \finer\stmt_2\OR\stmt[v]_2$.
-We conclude that it cannot be the case that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$ for any infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_1,\stmt[v]_2$.
-\end{proof}
-
-We next show that $\lequ$ is a pre-order, and that it is a partial order with respect to $\equ$-equivalence classes.
-
-\begin{prop}
-The relation $\lequ$ is reflexive and transitive.  
-\end{prop}
-\begin{proof}
-Reflexivity is immediate since for any $\stmt\in\tdomain$, we have $\stmt\finer\stmt$, so $\stmt\lequ\stmt$. (This works for both the new and old definition)
-
-Transitivity requires a little more work. Let $\stmt_1,\stmt_2,\stmt_3\in\tdomain$ with $\stmt_1\lequ\stmt_2\lequ\stmt_3$.
-
-{\bf Using the original definition:}
-We have several cases to consider.
-
-\emph{Case 1:} $\stmt_1\finer\stmt_2\finer\stmt_3$. Then $\stmt_1\finer\stmt_3$ and so $\stmt_1\lequ\stmt_3$.
-
-\emph{Case 2:} $\stmt_1\equ\stmt_2\equ\stmt_3$. Then $\stmt_1\equ\stmt_3$ (since $\equ$ is an equivalence relation), and so $\stmt_1\lequ\stmt_3$.
-
-\emph{Case 3:} $\stmt_1\equ\stmt_2\finer\stmt_3$. Then for some infinitesimals $\stmt[v]_1$ and $\stmt[v]_2$ (with respect to $\stmt[u]$), we have:
-$$
-\stmt_1\finer\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2\finer\stmt_3\OR\stmt[v]_2
-$$
-Unclear what to do here -- need $\stmt_3$ alone on the right-hand side somehow. Case 4 has the same problem. Let's try it with the modified definition (in the ``warning").
-
-{\bf Using the new definition:}
-Let $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ with $\stmt_1'\finer\stmt_2'$, and let $\stmt_2''\equ\stmt_2$ and $\stmt_3''\equ\stmt_3$ with $\stmt_2''\finer\stmt_3''$. 
-Then for an appropriate choice of infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_i,\stmt[w]_i,\stmt[v]_i',\stmt[w]_i'$, $i=1,2,3$, we have:
-$$
-\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_1'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \finer \stmt_2'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_2''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \finer
-$$
-$$
-\stmt_3''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \eqgran
-\stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1'
-$$
-Since $\stmt_1 \equ \stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3''$ and $\stmt_3 \equ \stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1' $, we have that $\stmt_1\lequ\stmt_3$.
-
-\end{proof}
-
-\begin{prop}
-If $\stmt_1\lequ\stmt_2$ and $\stmt_1\gequ \stmt_2$, then $\stmt_1\equ\stmt_2$.
-\end{prop}
-\begin{proof}
-With the original definition, this is immediate, since we would either have $\stmt_1\eqgran\stmt_2$ (which implies $\stmt_1\equ\stmt_2$), or we would already have $\stmt_1\equ\stmt_2$.
-
-With the modified definition, we need a bit more work. 
-\end{proof}
-
-Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotone?
-
-% case: possibilities in u: (1) finite: impose counting measure (2) countable: finitely additive but not countably (3) uncountable: countably additive, legitimate measure
-
-\subsection{Collapsing Orders}
-
-Searching around did not yield the theorems desired, so we will discuss the tools needed here.
-It turns out that it is not so straightforward to collapse a poset by an equivalence relation.
-Let's start with a possible definition:
-
-If $(P,\leq)$ is a poset and $\sim$ is a collapsing equivalence relation, then endow the set $P/\sim$ with the relation $\leq/\sim$ defined by $x/\sim \leq/\sim y/\sim$ whenever there exist $x'\in x/\sim$ and $y'\in y/\sim$ with $x'\leq y'$.
-
-It turns out that this is not sufficient. Here is an important example. Let our set be $\{w,x,y,z\}$ and a poset relation defined by $x\leq w$, $y\leq w$, $y\leq z$, with the following Hasse diagram:
-
-\begin{tikzpicture}
-    % First, locate each of the nodes and name them
-    \node (top) at (0,0) {$w$};
-    \node [below left  of=top] (left)  {$x$};
-    \node [below right of=top] (bottom right) {$y$};
-    \node [above right of=bottom right] (right) {$z$};
-
-    % Now draw the lines:
-    \draw [shorten <=-2pt, shorten >=-2pt] (top) -- (left);
-    \draw [shorten <=-2pt, shorten >=-2pt] (top) -- (bottom right);
-    \draw [shorten <=-2pt, shorten >=-2pt] (bottom right) -- (right);
-\end{tikzpicture}
-
-Let's consider two possible cases.
-Suppose first $w\sim y$, and $x$ and $z$ are only $\sim$ with themselves.
-Then in order to collapse $\leq$ under $\sim$, we will need $x \leq w\sim y \leq z$ for transitivity.
-
-Consider next the case that $w\sim y$ and $x\sim z$. In order to collapse this, we will need that $w\sim x\sim y \sim z$.
-
-These examples show that not every equivalence relation can collapse a poset into a poset (even if it only collapses a chain).
-
-Note that neither of the above problems are possible with {\bf totality}.
-It seems that what makes the most sense is collapsing by some equivalence relation, and then extending both the ordering and the equivalence relation to generate a legitimate poset.
-We can explore this later if needed.
-
-\section{Possible example}
-
-Let $\mathbb{N}$ be the set of natural numbers.
-We define the sets $D_{i,k} = \{ x \in \mathbb{N} \, | \, x \mod i = k \}$ with $k < i$.
-Therefore $D_{1,0} = \mathbb{N}$, $D_{2,0} = \{ 2, 4, 6, ...
-\}$, $D_{2,1} = \{ 1, 3, 5, ...
-\}$ and so on.
-Additionally, we say $D_{i,j} \finer D_{k,l}$ for all $k \leq i$ and for all $j$ and $l$.
-As needed, we may overload the notation by letting $D_{i,j} =  D_{i, j \mod i}$ where $j\mod i$ is the smallest positive residue of $j$ under $i$.
-
-Let $D_{i,j}$ and $D_{k,l}$ be two of these subsets.
-Consider the intersection $D_{i,j}\cap D_{k,l}$ and suppose it is nonempty; let $n$ be in this intersection.
-Then $D_{i,j} = D_{i,n}$ and $D_{k,l} = D_{k,n}$, so the intersection is given by $D_{i,n}\cap D_{j,n} = \{x\in \mathbb{N} \, | \, x \mod \gcd(i,j) = n\} = D_{\gcd(i,j),n}$.
-
-Note also that $D_{i,j}^C = \cup_{k=1,k\neq j}^{i-1}D_{i,k}$.
-
-It follows that any finite combination of unions, intersections, and complements of sets of the form $D_{i,j}$ will be empty, or countably infinite.
-
-Next, let $n\in\mathbb{N}$.
-Consider the union $S_n = \cup_{i=n+1}^{\infty}D_{i,0} = \{n+1,n+2,\ldots\}$, which may be generated by a countable union.
-Thus $S_n^C = \{1,2,\ldots,n\}$.
-By combining complements and intersections, we can obtain $\{n\} = S_{n-1}\cap S_n^C$.
-We can then generate arbitrary subsets of $\mathbb{N}$ using countable operations of union, intersection, and complement on the sets $D_{i,j}$.
-
-Claim: all finite operations of union, intersection, and complement of $D_{i,j}$'s yield sets that may be written as finite unions of $D_{i,j}$'s.
-Countable operations yield arbitrary subsets of $\mathbb{N}$.
-
-Possible source to read from: http://www.stat.cmu.edu/tr/tr814/tr814.pdf
-Note in abstract, gives $1/m$ probability (measure) to $D_{m,j}$ for each $j$. 
-\section{Other notes}
-
-Fineness may also capture the concept of physical dimension.
-In fact, two description of the same units are ``finitely comparable'' in the sense that one gives a finer description of the other of a finite factor.
-Descriptions of different unites are either ``infinitely comparable'' (i.e.
-areas are always bigger than lengths) or not comparable (i.e.
-position and momentum).
-Take phase space, points should be comparables and in fact should be equigranular $\eqgran$ so that we can compare sets of finite points.
-Areas are also comparable to each other, and are comparable to points.
-However, vertical lines (i.e.
-ranges in momentum alone) are not comparable to horizontal lines (i.e.
-ranges in position alone).
-Symplectic geometry, in fact, gives a size to areas and not to lines.
-
 \bibliographystyle{plain}

### 2021-03-26 Overviews/Informational granularity/main.tex
More cleanup for Informational granularity

@@
 
-\begin{prop}
+\begin{prop}\label{infinitesimal_disjunction}
 	The disjunction of two infinitesimals with respect to $\stmt[u]$ that are comparable\footnote{Is comparability required?
@@
 \end{proof}
+
+\subsection{Comparision and equivalence at $\stmt[u]$ scale}
+
+The general idea is to collapse the preorder given by fineness $\finer$ so that statements that are different up to an infinitesimal are equal.
+
+\def\equivu{\cong_{\stmt[u]}}
+
+\begin{defn}
+	Let $\stmt_1, \stmt_2 \in \tdomain$.
+We say $\stmt_1$ is \textbf{equivalent to $\stmt_2$ at $\stmt[u]$ scale}, noted $\stmt_1 \equ \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$.
+\end{defn}
+
+\begin{conj}
+	Equivalence at $\stmt[u]$ scale is an equivalence class.
+\end{conj}
+
+\begin{remark}
+	For reflexivity, let $\stmt \in \tdomain$.
+We have $\stmt \OR \impossibility \equiv \stmt \OR \impossibility$.
+Since $\impossibility$ is an infinitesimal for all $\stmt[u]$, $\stmt \equ \stmt$.
+For symmetry, note that the definition itself is symmetrical.
+
+	Transitivity is unclear.
+Let $\stmt_1 \equ \stmt_2 \equ \stmt_3$.
+Then we can write $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$ and $\stmt_2 \OR \stmt[w]_2 \eqgran \stmt_3 \OR \stmt[w]_3$.
+If the infinitesimal where incompatible we would have $\stmt_1 \OR \stmt[v]_1 \OR \stmt[w]_2 \eqgran \stmt_2 \OR \stmt[v]_2 \OR \stmt[w]_2 \eqgran \stmt_3 \OR \stmt[w]_3 \OR \stmt[v]_2$.
+If the infinitesimals were comparable, by proposition \ref{infinitesimal_disjunction} we would have that $\stmt[v]_1 \OR \stmt[w]_2$ and $\stmt[w]_3 \OR \stmt[v]_2$ are infinitesimals and therefore $\stmt_1 \equ \stmt_3$.
+
+	If we assume a uniform domain, we would have totality and we may also be able to show that incompatible infinitesimals can always be found.
+It may be that totality is not needed for proposition \ref{infinitesimal_disjunction}.
+It may be that we need to amend the definition of infinitesimals, such that the disjunction of two infinitesimals is an infinitesimals.
+We could also force the transitivity, even if the disjunction of two infinitesimals is not an infinitesimals.
+This would be like saying that, yes, in principles the assertions correspond to different granularity levels, but we are not to distinguish between them because the different is effectively zero.
+Understanding which options is best requires more work.
+\end{remark}
+
+\begin{conj}
+	Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$, then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
+\end{conj}
+
+\begin{remark}
+	What we'd like to have is that equivalence at $\stmt[u]$ scale merges contiguous granularity levels.
+It is yet to be understood what are the necessary preconditions for which this happens.
+One conjecture is that a uniform domain would be sufficient, though probably not necessary.
+\end{remark}
+
+[TODO: is there anything to be salvaged between here]
+
+We start with a simpler case:
+
+\begin{prop}
+	Let $\stmt_1,\stmt,\stmt[v]_1\in\tdomain$ such that $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$, and suppose $\stmt[v]_1$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt\equ\stmt_1$. 
+\end{prop}
+\begin{proof}[Using Prop \ref{uniformconsequence}]
+	If both of the $\finer$'s are actually equigranular, there is nothing to show.
+Next consider the case that $\stmt_1\sfiner\stmt\finer\stmt_1\OR\stmt[v]_1$.
+Then by Proposition \ref{uniformconsequence} there exists $\stmt_2$ such that $\stmt_2\narrower\stmt$ and $\stmt_2\eqgran\stmt_1$.
+But how do we put this together??
+What about the even simpler version of the lemma with $\narrower$ instead of $\finer$?
+Just use $\stmt\OR (\stmt_1\OR\stmt[v]_1 \AND \NOT\stmt)$.
+\end{proof}
+
+This seems problematic already:
+
+We have $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$.
+We need to find some $\stmt[w],\stmt[w]_1$ infinitesimal with respect to $\stmt[u]$, such that $\stmt\OR\stmt[w]\eqgran\stmt_1\OR\stmt[w]_1$.
+Since we don't know anything about the overlap of $\stmt$ and $\stmt_1$, we can't use any operations like $\stmt\AND\NOT\stmt_1$ to find something small.
+We need to know that additional statements exist in order to find $\stmt[w]$ and $\stmt[w]_1$.
+
+For example, what if every infinitesimal statement was narrower $(\narrower)$ than $\stmt$?
+In this case, we could have something like $\stmt_1\sfiner\stmt\sfiner\stmt_1\OR\stmt[x]$ for all statements (including all infinitesimals) $\stmt[x]$.
+It would then not be the case that $\stmt\equ\stmt_1$.
+
+\begin{prop}[Need something like this but not clear if correct!]
+	Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$. Then there exist statements  $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$, infinitesimal in $\stmt[u]$, such that:
+	$$
+	\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
+	$$
+\end{prop}
+
+Note that the existence of the right combination of incompatible statements is sufficient.
+However, that is stronger than we need and may be harder to prove.
+Problem: we have very few propositions about existence of statements!
+It is hard to ensure certain ones exist.
+
+Consider the following (extreme) case. Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$, but the infinitesimals which realize these relationships are $\stmt[v]_1$ and $\stmt[v]_2$, as follows:
+$$
+\stmt_1\OR\stmt[v]_1 \eqgran \stmt_2\OR\stmt[v]_2,\ \stmt_2\OR\stmt[v]_1\eqgran\stmt_3\OR\stmt[v]_2
+$$
+Without further axioms, it may not be possible to put this together to get transitivity (i.e. to find that $\stmt_1\equ\stmt_3$).
+
+\begin{prop}
+	The relation $\equ$ is an equivalence relation.
+\end{prop}
+\begin{proof}
+	Reflexivity and symmetry are immediate since $\eqgran$ is an equivalence relation.
+For transitivity, suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$.
+Let $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$ be infinitesimals with respect to $\stmt[u]$, such that $\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2$ and $\stmt_2\OR\stmt[w]_2\eqgran\stmt_3\OR\stmt[w]_3$.
+Then we have:
+	$$
+	\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
+	$$
+	Since $\stmt[v]_1\OR\stmt[w]_2$ and $\stmt[w]_3\OR\stmt[v]_2$ are infinitesimal with respect to $\stmt[u]$ (being finite disjuncts of infinitesimals), we have $\stmt_1\equ\stmt_3$. 
+\end{proof}
+
+[TODO: and here?]
+
+\begin{defn}\label{smalleru}
+	Let $\stmt_1, \stmt_2 \in \tdomain$.
+Then $\stmt_1$ is \textbf{smaller than} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \lequ \stmt_2$, if $\stmt_1 \finer \stmt_2$ or $\stmt_1 \equ \stmt_2$.
+We define $\stmt_1\leu\stmt_2$ by $\stmt_1\lequ\stmt_2$ and not $\stmt_1\equ\stmt_2$.
+\end{defn}
+
+\begin{defn}\label{smalleru2}
+	Let $\stmt_1, \stmt_2 \in \tdomain$.
+Then $\stmt_1$ is \textbf{smaller than} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \lequ \stmt_2$, if there exist statements $\stmt_1'$ and $\stmt_2'$ with $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ such that $\stmt_1'\finer\stmt_2'$.
+\end{defn}
+
+\begin{conj}
+	The two definitions are equivalent.
+\end{conj}
+
+\begin{remark}
+	The idea is that smallness in $\stmt[u]$ compares the equivalence classes at $\stmt[u]$ scale.
+In other words, we should be able to show that the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes is well-defined and monotonic.
+There may be a more natural order theoretic definition that automatically does this, and shows that smallness in $\stmt[u]$ is a preorder.
+\end{remark}
+
+[TODO: would remove this]
+
+{\bf Warning:} Transitivity doesn't seem to work as stated (see the proof attempt below).
+We propose the following modification: say $\stmt_1\lequ\stmt_2$ if there exist statements $\stmt_1'$ and $\stmt_2'$ with $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ such that $\stmt_1'\finer\stmt_2'$.
+Note that {\bf by totality}, it is sufficient to just say $\stmt_1\finer\stmt_2'$ (no need for $\stmt_1'$).
+
+[TODO: proof not clear]
+
+\begin{prop}
+	$\stmt_1 \leu \stmt_2$ if for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$. 
+\end{prop}
+\begin{proof}
+	Since $\impossibility$ is infinitesimal, we have $\stmt_1 = \stmt_1\OR\impossibility \sfiner \stmt_2$, and so by Definition \ref{smalleru}, we have that $\stmt_1\lequ\stmt_2$.
+Now, we must show it is not the case that $\stmt_1\equ \stmt_2$.
+
+	Let $\stmt[v]_1,\stmt[v]_2\in\tdomain$ be infinitesimals.
+Then $\stmt_1\OR\stmt[v]_1 \sfiner \stmt_2 \finer\stmt_2\OR\stmt[v]_2$.
+We conclude that it cannot be the case that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$ for any infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_1,\stmt[v]_2$.
+\end{proof}
+
+We next show that $\lequ$ is a pre-order, and that it is a partial order with respect to $\equ$-equivalence classes.
+
+\begin{conj}
+	The relation $\lequ$ is reflexive and transitive and therefore a preorder.
+\end{conj}
+
+[TODO: not clear that this works... don't we have to find infinitesimals that do not overlap?]
+\begin{proof}
+	Reflexivity is immediate since for any $\stmt\in\tdomain$, we have $\stmt\finer\stmt$, so $\stmt\lequ\stmt$. (This works for both the new and old definition)
+
+	Transitivity requires a little more work. Let $\stmt_1,\stmt_2,\stmt_3\in\tdomain$ with $\stmt_1\lequ\stmt_2\lequ\stmt_3$.
+
+	{\bf Using the original definition:}
+	We have several cases to consider.
+
+	\emph{Case 1:} $\stmt_1\finer\stmt_2\finer\stmt_3$. Then $\stmt_1\finer\stmt_3$ and so $\stmt_1\lequ\stmt_3$.
+
+	\emph{Case 2:} $\stmt_1\equ\stmt_2\equ\stmt_3$. Then $\stmt_1\equ\stmt_3$ (since $\equ$ is an equivalence relation), and so $\stmt_1\lequ\stmt_3$.
+
+	\emph{Case 3:} $\stmt_1\equ\stmt_2\finer\stmt_3$. Then for some infinitesimals $\stmt[v]_1$ and $\stmt[v]_2$ (with respect to $\stmt[u]$), we have:
+	$$
+	\stmt_1\finer\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2\finer\stmt_3\OR\stmt[v]_2
+	$$
+	Unclear what to do here -- need $\stmt_3$ alone on the right-hand side somehow. Case 4 has the same problem. Let's try it with the modified definition (in the ``warning").
+
+	{\bf Using the new definition:}
+	Let $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ with $\stmt_1'\finer\stmt_2'$, and let $\stmt_2''\equ\stmt_2$ and $\stmt_3''\equ\stmt_3$ with $\stmt_2''\finer\stmt_3''$. 
+	Then for an appropriate choice of infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_i,\stmt[w]_i,\stmt[v]_i',\stmt[w]_i'$, $i=1,2,3$, we have:
+	$$
+	\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_1'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \finer \stmt_2'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_2''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \finer
+	$$
+	$$
+	\stmt_3''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \eqgran
+	\stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1'
+	$$
+	Since $\stmt_1 \equ \stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3''$ and $\stmt_3 \equ \stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1' $, we have that $\stmt_1\lequ\stmt_3$.
+
+\end{proof}
+
+\begin{prop}
+	If $\stmt_1\lequ\stmt_2$ and $\stmt_1\gequ \stmt_2$, then $\stmt_1\equ\stmt_2$.
+\end{prop}
+\begin{proof}
+	This is immediate from the first definition, since we would either have $\stmt_1\eqgran\stmt_2$ (which implies $\stmt_1\equ\stmt_2$), or we would already have $\stmt_1\equ\stmt_2$.
+\end{proof}
+
+\begin{remark}
+	content...
+\end{remark}
+
+Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotone?
+
+% case: possibilities in u: (1) finite: impose counting measure (2) countable: finitely additive but not countably (3) uncountable: countably additive, legitimate measure
+
+\subsubsection{Collapsing Orders}
+[TODO: didn't we find a reference for this?]
+
+Searching around did not yield the theorems desired, so we will discuss the tools needed here.
+It turns out that it is not so straightforward to collapse a poset by an equivalence relation.
+Let's start with a possible definition:
+
+If $(P,\leq)$ is a poset and $\sim$ is a collapsing equivalence relation, then endow the set $P/\sim$ with the relation $\leq/\sim$ defined by $x/\sim \leq/\sim y/\sim$ whenever there exist $x'\in x/\sim$ and $y'\in y/\sim$ with $x'\leq y'$.
+
+It turns out that this is not sufficient. Here is an important example. Let our set be $\{w,x,y,z\}$ and a poset relation defined by $x\leq w$, $y\leq w$, $y\leq z$, with the following Hasse diagram:
+
+\begin{tikzpicture}
+% First, locate each of the nodes and name them
+\node (top) at (0,0) {$w$};
+\node [below left  of=top] (left)  {$x$};
+\node [below right of=top] (bottom right) {$y$};
+\node [above right of=bottom right] (right) {$z$};
+
+% Now draw the lines:
+\draw [shorten <=-2pt, shorten >=-2pt] (top) -- (left);
+\draw [shorten <=-2pt, shorten >=-2pt] (top) -- (bottom right);
+\draw [shorten <=-2pt, shorten >=-2pt] (bottom right) -- (right);
+\end{tikzpicture}
+
+Let's consider two possible cases.
+Suppose first $w\sim y$, and $x$ and $z$ are only $\sim$ with themselves.
+Then in order to collapse $\leq$ under $\sim$, we will need $x \leq w\sim y \leq z$ for transitivity.
+
+Consider next the case that $w\sim y$ and $x\sim z$. In order to collapse this, we will need that $w\sim x\sim y \sim z$.
+
+These examples show that not every equivalence relation can collapse a poset into a poset (even if it only collapses a chain).
+
+Note that neither of the above problems are possible with {\bf totality}.
+It seems that what makes the most sense is collapsing by some equivalence relation, and then extending both the ordering and the equivalence relation to generate a legitimate poset.
+We can explore this later if needed.
 
@@
 
-\subsection{Infinitesimals}
-
-\subsubsection{Equivalence at $\stmt[u]$ scale}
-
-\def\equivu{\cong_{\stmt[u]}}
-
-\begin{defn}
-Let $\stmt_1, \stmt_2 \in \tdomain$.
-We say $\stmt_1$ is \textbf{equivalent to $\stmt_2$ at $\stmt[u]$ scale}, noted $\stmt_1 \equivu \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$.
-\end{defn}
-
-\begin{prop}
-Equivalence at $\stmt[u]$ scale is an equivalence class.
-\end{prop}
-\begin{proof}
-For reflexivity, let $\stmt \in \tdomain$.
-We have $\stmt \OR \impossibility \equiv \stmt \OR \impossibility$.
-Since $\impossibility$ is an infinitesimal for all $\stmt[u]$ (TODO: add proposition), $\stmt \equivu \stmt$.
-For symmetry, note that the definition itself is symmetrical.
-For transitivity, let $\stmt_1 \equivu \stmt_2 \equivu \stmt_3$.
-Then we have $\stmt_1 \OR \stmt[v]_1 \equiv \stmt_2 \OR \stmt[v]_2$ and $\stmt_2 \OR \stmt[w]_2 \equiv \stmt_3 \OR \stmt[w]_3$.
-\end{proof}
-
-\subsubsection{Greatness in $\stmt[u]$}
-
-First define ``as big in $\stmt[u]$'' $\equ$.
-
-\begin{defn}
-Let $\stmt_1, \stmt_2 \in \tdomain$.
-Then $\stmt_1$ is \textbf{as big as} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \equ \stmt_2$, if we can find $\stmt[v]_1, \stmt[v]_2 \in \tdomain$ that are infinitesimal with respect to $\stmt[u]$ such that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$.
-\end{defn}
-
-\begin{prop}
-Suppose $\tdomain$ is uniform.
-Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$, then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
-\end{prop}
-
-We start with a simpler case:
-
-\begin{prop}
-Let $\stmt_1,\stmt,\stmt[v]_1\in\tdomain$ such that $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$, and suppose $\stmt[v]_1$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt\equ\stmt_1$. 
-\end{prop}
-\begin{proof}[Using Prop \ref{uniformconsequence}]
-If both of the $\finer$'s are actually equigranular, there is nothing to show.
-Next consider the case that $\stmt_1\sfiner\stmt\finer\stmt_1\OR\stmt[v]_1$.
-Then by Proposition \ref{uniformconsequence} there exists $\stmt_2$ such that $\stmt_2\narrower\stmt$ and $\stmt_2\eqgran\stmt_1$.
-But how do we put this together??
-What about the even simpler version of the lemma with $\narrower$ instead of $\finer$?
-Just use $\stmt\OR (\stmt_1\OR\stmt[v]_1 \AND \NOT\stmt)$.
-\end{proof}
-
-This seems problematic already:
-
-We have $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$.
-We need to find some $\stmt[w],\stmt[w]_1$ infinitesimal with respect to $\stmt[u]$, such that $\stmt\OR\stmt[w]\eqgran\stmt_1\OR\stmt[w]_1$.
-Since we don't know anything about the overlap of $\stmt$ and $\stmt_1$, we can't use any operations like $\stmt\AND\NOT\stmt_1$ to find something small.
-We need to know that additional statements exist in order to find $\stmt[w]$ and $\stmt[w]_1$.
-
-For example, what if every infinitesimal statement was narrower $(\narrower)$ than $\stmt$?
-In this case, we could have something like $\stmt_1\sfiner\stmt\sfiner\stmt_1\OR\stmt[x]$ for all statements (including all infinitesimals) $\stmt[x]$.
-It would then not be the case that $\stmt\equ\stmt_1$.
-
-\begin{prop}[Need something like this but not clear if correct!]
-Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$. Then there exist statements  $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$, infinitesimal in $\stmt[u]$, such that:
-$$
-\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
-$$
-\end{prop}
-
-Note that the existence of the right combination of incompatible statements is sufficient.
-However, that is stronger than we need and may be harder to prove.
-Problem: we have very few propositions about existence of statements!
-It is hard to ensure certain ones exist.
-
-Consider the following (extreme) case. Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$, but the infinitesimals which realize these relationships are $\stmt[v]_1$ and $\stmt[v]_2$, as follows:
-$$
-\stmt_1\OR\stmt[v]_1 \eqgran \stmt_2\OR\stmt[v]_2,\ \stmt_2\OR\stmt[v]_1\eqgran\stmt_3\OR\stmt[v]_2
-$$
-Without further axioms, it may not be possible to put this together to get transitivity (i.e. to find that $\stmt_1\equ\stmt_3$).
-
-\begin{prop}
-The relation $\equ$ is an equivalence relation.
-\end{prop}
-\begin{proof}
-Reflexivity and symmetry are immediate since $\eqgran$ is an equivalence relation.
-For transitivity, suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$.
-Let $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$ be infinitesimals with respect to $\stmt[u]$, such that $\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2$ and $\stmt_2\OR\stmt[w]_2\eqgran\stmt_3\OR\stmt[w]_3$.
-Then we have:
-$$
-\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
-$$
-Since $\stmt[v]_1\OR\stmt[w]_2$ and $\stmt[w]_3\OR\stmt[v]_2$ are infinitesimal with respect to $\stmt[u]$ (being finite disjuncts of infinitesimals), we have $\stmt_1\equ\stmt_3$. 
-\end{proof}
-
-\begin{defn}\label{smalleru}
-Let $\stmt_1, \stmt_2 \in \tdomain$.
-Then $\stmt_1$ is \textbf{smaller than} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \lequ \stmt_2$, if $\stmt_1 \finer \stmt_2$ or $\stmt_1 \equ \stmt_2$.
-We define $\stmt_1\leu\stmt_2$ by $\stmt_1\lequ\stmt_2$ and not $\stmt_1\equ\stmt_2$.
-\end{defn}
-{\bf Warning:} Transitivity doesn't seem to work as stated (see the proof attempt below).
-We propose the following modification: say $\stmt_1\lequ\stmt_2$ if there exist statements $\stmt_1'$ and $\stmt_2'$ with $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ such that $\stmt_1'\finer\stmt_2'$.
-Note that {\bf by totality}, it is sufficient to just say $\stmt_1\finer\stmt_2'$ (no need for $\stmt_1'$).
-
- \begin{prop}
-$\stmt_1 \leu \stmt_2$ if for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$. 
-\end{prop}
-\begin{proof}
-Since $\impossibility$ is infinitesimal, we have $\stmt_1 = \stmt_1\OR\impossibility \sfiner \stmt_2$, and so by Definition \ref{smalleru}, we have that $\stmt_1\lequ\stmt_2$.
-Now, we must show it is not the case that $\stmt_1\equ \stmt_2$.
-
-Let $\stmt[v]_1,\stmt[v]_2\in\tdomain$ be infinitesimals.
-Then $\stmt_1\OR\stmt[v]_1 \sfiner \stmt_2 \finer\stmt_2\OR\stmt[v]_2$.
-We conclude that it cannot be the case that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$ for any infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_1,\stmt[v]_2$.
-\end{proof}
-
-We next show that $\lequ$ is a pre-order, and that it is a partial order with respect to $\equ$-equivalence classes.
-
-\begin{prop}
-The relation $\lequ$ is reflexive and transitive.  
-\end{prop}
-\begin{proof}
-Reflexivity is immediate since for any $\stmt\in\tdomain$, we have $\stmt\finer\stmt$, so $\stmt\lequ\stmt$. (This works for both the new and old definition)
-
-Transitivity requires a little more work. Let $\stmt_1,\stmt_2,\stmt_3\in\tdomain$ with $\stmt_1\lequ\stmt_2\lequ\stmt_3$.
-
-{\bf Using the original definition:}
-We have several cases to consider.
-
-\emph{Case 1:} $\stmt_1\finer\stmt_2\finer\stmt_3$. Then $\stmt_1\finer\stmt_3$ and so $\stmt_1\lequ\stmt_3$.
-
-\emph{Case 2:} $\stmt_1\equ\stmt_2\equ\stmt_3$. Then $\stmt_1\equ\stmt_3$ (since $\equ$ is an equivalence relation), and so $\stmt_1\lequ\stmt_3$.
-
-\emph{Case 3:} $\stmt_1\equ\stmt_2\finer\stmt_3$. Then for some infinitesimals $\stmt[v]_1$ and $\stmt[v]_2$ (with respect to $\stmt[u]$), we have:
-$$
-\stmt_1\finer\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2\finer\stmt_3\OR\stmt[v]_2
-$$
-Unclear what to do here -- need $\stmt_3$ alone on the right-hand side somehow. Case 4 has the same problem. Let's try it with the modified definition (in the ``warning").
-
-{\bf Using the new definition:}
-Let $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ with $\stmt_1'\finer\stmt_2'$, and let $\stmt_2''\equ\stmt_2$ and $\stmt_3''\equ\stmt_3$ with $\stmt_2''\finer\stmt_3''$. 
-Then for an appropriate choice of infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_i,\stmt[w]_i,\stmt[v]_i',\stmt[w]_i'$, $i=1,2,3$, we have:
-$$
-\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_1'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \finer \stmt_2'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_2''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \finer
-$$
-$$
-\stmt_3''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \eqgran
-\stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1'
-$$
-Since $\stmt_1 \equ \stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3''$ and $\stmt_3 \equ \stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1' $, we have that $\stmt_1\lequ\stmt_3$.
-
-\end{proof}
-
-\begin{prop}
-If $\stmt_1\lequ\stmt_2$ and $\stmt_1\gequ \stmt_2$, then $\stmt_1\equ\stmt_2$.
-\end{prop}
-\begin{proof}
-With the original definition, this is immediate, since we would either have $\stmt_1\eqgran\stmt_2$ (which implies $\stmt_1\equ\stmt_2$), or we would already have $\stmt_1\equ\stmt_2$.
-
-With the modified definition, we need a bit more work. 
-\end{proof}
-
-Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotone?
-
-% case: possibilities in u: (1) finite: impose counting measure (2) countable: finitely additive but not countably (3) uncountable: countably additive, legitimate measure
-
-\subsection{Collapsing Orders}
-
-Searching around did not yield the theorems desired, so we will discuss the tools needed here.
-It turns out that it is not so straightforward to collapse a poset by an equivalence relation.
-Let's start with a possible definition:
-
-If $(P,\leq)$ is a poset and $\sim$ is a collapsing equivalence relation, then endow the set $P/\sim$ with the relation $\leq/\sim$ defined by $x/\sim \leq/\sim y/\sim$ whenever there exist $x'\in x/\sim$ and $y'\in y/\sim$ with $x'\leq y'$.
-
-It turns out that this is not sufficient. Here is an important example. Let our set be $\{w,x,y,z\}$ and a poset relation defined by $x\leq w$, $y\leq w$, $y\leq z$, with the following Hasse diagram:
-
-\begin{tikzpicture}
-    % First, locate each of the nodes and name them
-    \node (top) at (0,0) {$w$};
-    \node [below left  of=top] (left)  {$x$};
-    \node [below right of=top] (bottom right) {$y$};
-    \node [above right of=bottom right] (right) {$z$};
-
-    % Now draw the lines:
-    \draw [shorten <=-2pt, shorten >=-2pt] (top) -- (left);
-    \draw [shorten <=-2pt, shorten >=-2pt] (top) -- (bottom right);
-    \draw [shorten <=-2pt, shorten >=-2pt] (bottom right) -- (right);
-\end{tikzpicture}
-
-Let's consider two possible cases.
-Suppose first $w\sim y$, and $x$ and $z$ are only $\sim$ with themselves.
-Then in order to collapse $\leq$ under $\sim$, we will need $x \leq w\sim y \leq z$ for transitivity.
-
-Consider next the case that $w\sim y$ and $x\sim z$. In order to collapse this, we will need that $w\sim x\sim y \sim z$.
-
-These examples show that not every equivalence relation can collapse a poset into a poset (even if it only collapses a chain).
-
-Note that neither of the above problems are possible with {\bf totality}.
-It seems that what makes the most sense is collapsing by some equivalence relation, and then extending both the ordering and the equivalence relation to generate a legitimate poset.
-We can explore this later if needed.
-
-\section{Possible example}
+\section{Examples}
+
+Some ideas on distributions on natural numbers can be found in \cite{Schirokauer2007}.
+% Also available at http://www.stat.cmu.edu/tr/tr814/tr814.pdf
+
+\subsection{Distribution on natural numbers through mod classes}
 
@@
 
+\subsubsection{The $\sigma$-algebra generated by the sets is the powerset}
+
 Let $D_{i,j}$ and $D_{k,l}$ be two of these subsets.
@@
 
-Possible source to read from: http://www.stat.cmu.edu/tr/tr814/tr814.pdf
-Note in abstract, gives $1/m$ probability (measure) to $D_{m,j}$ for each $j$.
+\subsubsection{A finitely-additive measure on infinite sets}
+
+Take $\mathbb{N}$ as the unit.
+For any $i$, we have $D_{i,j} \finer D_{i,l} \finer D_{i,j}$.
+Therefore $D_{i,j} \eqgran D_{i,l}$ for any $j$ and $l$.
+We have $1 = \mu_{\mathbb{N}}(\mathbb{N}) = \mu_{\mathbb{N}}(\bigOR_{j=1}^i D_{i,j}) = \sum_{j=1}^i \mu_{\mathbb{N}}( D_{i,j}) = \sum_{j=1}^i \mu_{\mathbb{N}}( D_{i,0}) = i \mu_{\mathbb{N}}( D_{i,0}) $.
+Therefore $\mu_{\mathbb{N}}( D_{i,j}) = 1/i$ for any $i$ and $j$.
 
@@
 
+In this section we summarize ideas that did not work.
+
+\subsection{Transport of sequences}
+
+[TODO: example on how moving a sequence to one that is equigranular at all finite steps does not give the same limit]
+
 \subsection{Maximal chains}
+[TODO: boil this down to the essentials
 

### 2021-03-30 Overviews/Informational granularity/main.tex
InfGran: review with Christine until family of measures

@@
-\documentclass[11pt]{article}
+\documentclass[10pt, onecolumn, nofootinbib]{revtex4-1}
 \usepackage[utf8]{inputenc}
@@
 \input{theorems}
+\frenchspacing
 
@@
 
-\title{Assumptions of Physics blueprint: information granularity as a common foundation for geometry, measures and probability.}
+\begin{document}
+
+	%TODO: add e-mail
+	%TODO: not have page break on first page
+
+\title{Assumptions of Physics blueprint: information granularity as a common foundation for geometry, measures and probability}
 \author{Gabriele Carcassi, Mark Greenfield, Christine Aidala }
+\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
 \date{\today}
 
-\begin{document}
-
-\maketitle
-
 \begin{abstract}
-    This paper presents ideas and progress towards an order theoretic framework that can be used as a common foundation for geometrical, measure theoretic and probabilistic concepts in the sciences.
-This would address foundational questions such as: under what assumptions standard tools like calculus are physically meaningful?
-How do geometrical structures arise in physics?
+    This paper presents ideas and progress towards an order theoretic framework that can be used as a common foundation for geometric, measure theoretic and probabilistic concepts in the sciences.
+This would address foundational questions such as: under what assumptions are standard tools like calculus physically meaningful?
+How do geometric structures arise in physics?
 How can the idea of physical dimension be properly captured mathematically?
-The working theory is that geometrical and measure theoretic structures capture one and only one thing: how coarse or fine grained are the statements within a scientific theory.
-They should  be recovered from a single binary operator, that allows to compare two statements and say whether one provides a finer level of description than the other.
+The working hypothesis is that geometric and measure theoretic structures capture one and only one thing: how coarse or fine grained the statements are within a scientific theory.
+Those structures should be recovered from a single binary operator, which allows us to compare two statements and say whether one provides a finer level of description than the other.
 Mathematically, fineness imposes a preorder on the algebra of statements of a scientific theory.
-Measure theoretical and geometrical notions will be recovered by picking a unit and quantifying the degree of precision in terms of that unit: ``the mass of the object is between 1 and 3 Kg'' is twice less precise than ``the mass of the object is between 0 and 1 Kg''.
-The objective is to find necessary and/or sufficient conditions for such constructions which can be given a straight forward physical motivation.
+Measure theoretic and geometric notions will be recovered by picking a unit and quantifying the degree of precision in terms of that unit: ``the mass of the object is between 1 and 3 Kg'' is half as precise as ``the mass of the object is between 0 and 1 Kg''.
+The objective is to find necessary and/or sufficient conditions for such constructions which can be given a straightforward physical motivation.
 %    As in comparative probability, we start with a $\sigma$-complete Boolean algebra that represents all possible descriptions in a given context and impose a preorder.
@@
 
+\maketitle
+
 \def\eqgran{\doteq}
@@
 
-\emph{This paper presents the goal and the current status of our research. It is intended to gather early feedback, including pointers to other relevant work.}
-
-The overall goal of Assumptions of physics is to to identify a handful of physical principles from which the basic laws can be rigorously derived.
-We have already established that topologies and $\sigma$-algebras can be recovered from requirements of experimental verifiability.\cite{AoPBook,AoPOverviewVerifiability} The next step is to understand what are the underpinning of geometrical and measure theoretic structures (including probability) which are layered on top of topologies and $\sigma$-algebras.
+\emph{This paper presents the goal and the current status of our research. It is intended to gather early feedback, including pointers to other relevant work. Feedback is welcome and encouraged.}
+
+The overall goal of Assumptions of Physics is to identify a handful of physical principles from which the basic laws can be rigorously derived.
+We have already established that topologies and $\sigma$-algebras can be recovered from requirements of experimental verifiability.\cite{AoPBook,AoPOverviewVerifiability} See also (TODO: Kevin Kelly).
+The next step is to understand the underpinnings of geometric and measure theoretic structures (including probability) which are layered on top of topologies and $\sigma$-algebras.
 As usual, a formal construction from scratch will force us to make explicit the physical requirements that are baked into the use of the common concepts and provide us with a framework that is, hopefully, still valid when those assumptions fail.
-
-Note that often in physics geometric structures are recovered from other geometric structures through the idea of emergence.
-For example, substantial effort is put in constructing theories where space-time emerge from other more fundamental structures.
+%TODO: add topology proceeding, update book reference
+
+Note that often in physics, geometric structures are recovered from other geometric structures through the idea of emergence.
+For example, substantial effort is put into constructing theories where space-time emerges from other more fundamental structures.
 This does not solve the problem: it simply pushes it backwards.
@@
 
-The idea is that while experimental verifiability keeps track of which statements in the theory are experimentally verifiable, informational granularity keeps track of which provides a finer level of description.
-\textbf{The working theory is that this, and only this, is what geometrical and measure theoretic structure capture: how coarse or fine grained are the statements within a scientific theory.} Finding the fundamental axioms associated to the notion of information granularity is therefore the central problem.
+The idea is that while experimental verifiability keeps track of which statements in the theory are experimentally verifiable, informational granularity keeps track of which provide a finer level of description.
+\textbf{The working hypothesis is that this, and only this, is what geometric and measure theoretic structures capture: how coarse or fine grained the statements are within a scientific theory.} Finding the fundamental axioms associated to the notion of information granularity is therefore the central problem.
 
@@
 Therefore the notion of fineness is linked to the notion of ``counting'' of the possible outcomes.
-It should be no surprise, then, that the same concept is compatible with the different disciplines above: geometrical size quantifies the ways we can arrange objects; probability theory quantifies the ratio of expected cases; information theory needs to quantify the different possible messages stored, computed or transmitted through an information system.
+It should be no surprise, then, that the same concept is compatible with the different disciplines above: geometric size quantifies the ways we can arrange objects; probability theory quantifies the ratio of expected cases; information theory quantifies the different possible messages stored, computed or transmitted through an information system.
 Therefore the common structure we propose is not simply a mathematical pattern among disconnected structures: it is a single conceptual framework.
@@
 We found a few ideas from comparative probability useful, but our goal cannot be achieved with a single measure.
-A single measure can only compare objects with a finite measure.
+A single measure can only compare objects of finite measure.
 All objects with zero measure (or infinite measure) are indistinguishable.
@@
 Therefore we need to construct a family of measures, each identified by picking a particular statement that acts as a reference, as a unit.
-This provides a much richer structure, that is both interesting mathematically and more scientifically useful.
-
-As for geometry, geometric structures provide measures for different dimensionality by starting from a lowest definition (i.e.
-line distance for Riemannian and surface are for symplectic) and then constructing higher dimensional measures through products.
-While want to recover those cases as well, this is not general enough.
+This provides a much richer structure, which is both interesting mathematically and more scientifically useful.
+
+As for geometry, geometric structures provide measures for different dimensionality by starting from a lowest dimension (i.e.
+line distance for Riemannian geometry and surface area for symplectic geometry) and then constructing higher dimensional measures through products.
+While we want to recover those cases as well, this is not general enough.
 In physics there will be dimensions that are not directly comparable (e.g.
@@
 The full detail of why this is the correct abstraction is discussed in detail in our previous work.\cite{AoPBook} It is a property of this type of spaces that each statement $\stmt \in \tdomain$ can be expressed as a disjunction of possible cases, of the  \textbf{possibilities} of the domain.\footnote{For example, ``the mass of the object is less than 1 Kg'' is the disjunction (i.e.
-the logical OR) of all the statements of the form ``the mass of the object is exact $x$ Kg'' where $0 \leq x < 1$.} Formally, each possibility is a minterm\footnote{A minterm is a disjunction in which each term appears once, either negated or not.} of the countable basis.
+the logical OR) of all the statements of the form ``the mass of the object is exactly $x$ Kg'' where $0 \leq x < 1$.} Formally, each possibility is a minterm\footnote{A minterm is a disjunction in which each term appears once, either negated or not.} of the countable basis.
 The theoretical domain can then equivalently be represented by a $\sigma$-algebra.\footnote{This conceptually maps perfectly to what one has in probability: a sample space (i.e.
@@
 
-The whole program can be summarized in the following question: what intuitive conceptual properties does fineness have physically that, once properly formalized mathematically, allow us to recover measures, probability and geometrical structures?\footnote{In terms of the $\sigma$-algebra, how do we go from being able to compare the size of sets to being able to quantify it?}
+The whole program can be summarized in the following question: what intuitive conceptual properties does fineness have physically that, once properly formalized mathematically, allow us to recover measures, probability and geometric structures?\footnote{In terms of the $\sigma$-algebra, how do we go from being able to compare the size of sets to being able to quantify it?}
 
@@
 
-The first set of properties are the ones we must expect fineness to have in all cases.
+The first properties are the ones we must expect fineness to have in all cases.
 At this point, there are three, identified by DeFinetti\cite{definetti} for comparative probability, that we believe are necessary and that work as simple and intuitive starting points.
 
-\textbf{Boundness}.
+\textbf{Boundedness}.
 As we mentioned in the introduction, a description is finer than another if it can be true in fewer cases.
@@
 \textbf{Transitivity}.
-Additionally, if $\stmt_1$ gives finer description than $\stmt_2$ and $\stmt_2$ gives a finer description than $\stmt_3$ then it must be that $\stmt_1$ gives a finer description than $\stmt_3$.
+Additionally, if $\stmt_1$ gives a finer description than $\stmt_2$ and $\stmt_2$ gives a finer description than $\stmt_3$ then it must be that $\stmt_1$ gives a finer description than $\stmt_3$.
 This means that fineness is transitive, and this is the second required property.
@@
 Lastly, if we take two descriptions and we make them less precise ``in the same way'', we must have the same relationship in fineness.
-More precisely, suppose we have two descriptions $\stmt_1$ and $\stmt_2$ and we pick a third description $\stmt$ that is incompatible\footnote{Two statements are incompatible if they can't be both true at the same time: their conjunction $\stmt_1 \AND \stmt \equiv \impossibility$ is impossible.
-This is the analogue of disjoint sets in a $\sigma$-algebra.} with both.
-The respective disjunctions $\stmt_1 \OR \stmt$ and $\stmt_2 \OR \stmt$ will be respectively ``equally'' broader than the respective original statement.
+More precisely, suppose we have two descriptions $\stmt_1$ and $\stmt_2$ and we pick a third description $\stmt$ that is incompatible\footnote{Two statements are incompatible if they can't both be true at the same time: their conjunction $\stmt_1 \AND \stmt \equiv \impossibility$ is impossible.
+This is the analogue of disjoint sets in a $\sigma$-algebra.} with both (noted $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$).
+The respective disjunctions $\stmt_1 \OR \stmt$ and $\stmt_2 \OR \stmt$ will be ``equally'' broader than the original statements.
 Therefore we will want them to be ``equally'' coarser as well.
@@
 \begin{description}
-    \item[Preorder.] Fineness is a preorder over the theoretical domain $\tdomain$.
-    \item[Monotonicity.] Fineness respects narrowness: $\stmt_1 \narrower \stmt_2$ implies $\stmt_1 \finer \stmt_2$.
-    \item[Equigranularity is an equivalence.] Equigranularity, the symmetrization of fineness, is an equivalence class, and fineness is an order of those equivalence classes.
+    \item[Preorder] Fineness is a preorder over the theoretical domain $\tdomain$.
+    \item[Monotonicity] Fineness respects narrowness: $\stmt_1 \narrower \stmt_2$ implies $\stmt_1 \finer \stmt_2$.
+    \item[Equigranularity is an equivalence] Equigranularity $\eqgran$, the symmetrization of fineness, is an equivalence, and fineness is a partial order of its equivalence classes.
 That is, fineness orders the levels of description.
@@
 
-While we are not convinced these are sufficient properties to define fineness, we are also not found other properties that must be convincingly necessary. This is still an open area.
-
-\subsubsection{Optional properties for fineness}
-
-Optional properties for fineness include properties that are desirable, but are known to fail in some cases. These will be properties of a particular domain, rather than fineness itself.
+While we are not convinced these are sufficient properties to define fineness, we have not found other properties that we believe to be necessary. This is still an open area.
+
+\subsubsection{Domain-specific properties for fineness}
+
+Domain-specific properties for fineness include properties that are desirable, but are known to fail in some cases. These will be properties of a particular domain, rather than fineness itself.
 
@@
 Therefore it is natural to assume that all possibilities are equigranular.
-In fact, this is a necessary property if we want to have a unique measure when we take a possibility as a unit.\footnote{The measure, in this case, is simply the counting measure and equigranularity of the possibilities plays a role similar to the principle of indifference in classical probability} However, this property can be easily violated by creating a subdomain.
+In fact, this is a necessary property if we want to have a unique measure when we take a possibility as a unit.\footnote{The measure, in this case, is simply the counting measure and equigranularity of the possibilities plays a role similar to the principle of indifference in classical probability.} However, this property can be easily violated by creating a subdomain.
 
@@
 every Borel set of possibilities) needs to be comparable with every other if we are to assign a unique number.
-However, statements whose size is defined in units of different physical dimensions (i.e.
+However, statements whose size is defined in units of different physical dimensions (e.g.
 length, pressure, mass) should not be comparable.
@@
 For example, in phase space (i.e.
-symplectic manifold) we would expect points to be comparable and equigranular $\eqgran$: each state gives the same description.
+in a symplectic manifold) we would expect points to be comparable and equigranular: each state gives the same description.
 This means that we can compare sets of finite points.
-Finite areas are also comparable to each other (through integration of the symplectic form), and are comparable to points.
+Finite areas are also comparable to each other (through integration of the symplectic form), and are comparable to points: they will be infinitely coarser.
 However, vertical lines (i.e.
@@
 
-Another counter example come from infinite ranges.
+Another counterexample comes from infinite ranges.
 Take the real line and consider the set of all positive numbers and the set of all negative numbers: should they be comparable?
-If we say they are equal, then that would tell us that there are more cases before 0 than after.
-This means that the space is not symmetrical under translation.
+If we say they are equal, then that would tell us that there are the same number of cases before and after 0.
+This means that the space is not symmetric under translation.
 If the real values represent distance between points, this is bad as it violates the idea that space is uniform.
-If the real values represents charge, however, this would actually make sense: the idea that there are exactly the same values for positive and negative charges does make sense, and the symmetry we are imposing is natural since zero is really a special number.
+If the real values represent charge, however, this would actually make sense: the idea that there are exactly the same values for positive and negative charges does make sense, and the symmetry we are imposing is natural since zero is really a special number.
 Therefore different choices of comparability for the same Borel algebra is actually a feature we do not want to lose in general.
@@
 However, once we pick the unit, we should be able to find the ``right'' set of statements on which to create the measure.
-Roughly speaking, two description of the same units are ``finitely comparable'' in the sense that one gives a finer description of the other of a finite factor.
-Descriptions of different unites are either ``infinitely comparable'' (i.e.
-areas are always bigger than lengths) or not comparable (i.e.
+Roughly speaking, two descriptions of the same units are ``finitely comparable'' in the sense that one gives a finer description than the other by a finite factor.
+Descriptions of different units are either ``infinitely comparable'' (e.g.
+areas are always bigger than lengths) or not comparable (e.g.
 position and momentum).
-One approach would be to ask the ``largest set of statements mutually comparable in fineness that contain the unit''.
+One approach would be to find the ``largest set of statements mutually comparable in fineness that contain the unit''.
 In general, we have no expectation that this would be unique.
-So one question would be: when is this true?
+So one question would be: when is this set unique?
 
@@
 Physically this would make sense: we would be saying that given a set of statements $S \subseteq \tdomain$, we can always find a level of description that is the finest among all coarser descriptions than all the statements (i.e.
-$\bigveedot S$) and a level of description that is the coarsest among all finest descriptions than all the statements (i.e.
+$\bigveedot S$) and a level of description that is the coarsest among all finer descriptions than all the statements (i.e.
 $\bigwedgedot S$).
@@
 Let us note $\dot{\stmt} = {\stmt}_{/\eqgran}$ the level of granularity of $\stmt$.
-Now suppose that $\stmt_1 \eqgran \stmt_2 \nequiv \impossibility$ and $\stmt_1 \ncomp \stmt_2$.
+Now suppose that $\stmt_1 \eqgran \stmt_2 \nequiv \impossibility$ and $\stmt_1 \ncomp \stmt_2$ (recall $\ncomp$ means incompatible).
 Then $\dot{\stmt}_1 = \dot{\stmt}_2 \neq \dot{\impossibility}$.
@@
 We also need to understand how to define the addition and subtraction.
-Conceptually, the addition of two level of precision $\dot{\stmt}_1 + \dot{\stmt}_2$ means returning the level of precision that can be achieved by taking the disjunction $\stmt_1 \OR \stmt_2$ of two incompatible statements $\stmt_1 \in \dot{\stmt}_1$ and $\stmt_2 \in \dot{\stmt}_2$.
+Conceptually, the addition of two granularity levels $\dot{\stmt}_1 + \dot{\stmt}_2$ means returning the granularity level that can be achieved by taking the disjunction $\stmt_1 \OR \stmt_2$ of two incompatible statements $\stmt_1 \in \dot{\stmt}_1$ and $\stmt_2 \in \dot{\stmt}_2$.
 This, of course, is not possible if the statements are ``bigger'' than ``half'' the space (e.g.
 $\stmt_1 \eqgran \stmt_2$, $\stmt_1 \comp \stmt_2$ and $\stmt_1 \OR \stmt_2 \equiv \certainty$).
-A more order theoretical definition would be to take all possible disjuctions of $\stmt[v]_1 \OR \stmt[v]_2$ such that $\stmt[v]_1 \in \dot{\stmt}_1$ and $\stmt[v]_2 \in \dot{\stmt}_2$ and return the coarsest level of description.
+A more order theoretic definition would be to take all possible disjuctions of $\stmt[v]_1 \OR \stmt[v]_2$ such that $\stmt[v]_1 \in \dot{\stmt}_1$ and $\stmt[v]_2 \in \dot{\stmt}_2$ and return the coarsest level of description.
 That is, the sum is the least upper bound in fineness of all possible disjunctions.
@@
 In many of our proofs, we ran into an issue of not being able to guarantee existence of certain statements.
-Several of the definitions we wanted included statements like ``say x holds if there exist statement y such that w." However, we have no axioms that guarantee any existence of any statements beyond $\certainty$ and $\impossibility$.
+Several of the definitions we want have the form ``x holds if there exists statement y such that w." However, we have no axioms that guarantee the existence of any statements beyond $\certainty$ and $\impossibility$.
 In the case of an infinite collection of possibilities, we have no way to guarantee existence of statements of various levels of fineness when all statements in question are compatible with infinitely many possibilities.
@@
 We may also want to say something about existence of enough statements representing each equivalence class so that sum and product can be well-defined.
-For example, if $\stmt_1\finer\stmt_2$, we want to say that there exists $\stmt_1'$ with $\stmt_1'\eqgran\stmt_1$ and $\stmt_1'\narrower\stmt_2$, as well as $\stmt_1''$ such that $\stmt_2\AND\stmt_1''$ is minimal in some sense ($\stmt_2\ncomp\stmt_1''$ if they do not fill the space).
+For example, if $\stmt_1\finer\stmt_2$, we may want to say that there exists $\stmt_1'$ with $\stmt_1'\eqgran\stmt_1$ and $\stmt_1'\narrower\stmt_2$, as well as $\stmt_1''$ such that $\stmt_2\AND\stmt_1''$ is minimal in some sense ($\stmt_2\ncomp\stmt_1''$ if they do not fill the space).
 This is not at all guaranteed by the decomposition of statements into possibilities, since cardinality does not help us here.
 
-\textbf{Bigger and smaller halfs}.
+\textbf{Bigger and smaller ``halves''}.
 One thing that appears to be missing is the following.
 Consider the following domain with four possibilities $\{x_1, x_2, x_3, x_4\}$.
-Each statement is the disjunction of a set of possibilities and let us use the lower indexes to indicate them.
+Each statement is the disjunction of a set of possibilities and let us use the lower indices to indicate them.
 For example, $\stmt_{12} = x_1 \OR x_2$.
-Now, suppose $\stmt_{12} \broader \stmt_{34}$ and $\stmt_{14} \broader \stmt_{23}$.
+Now, suppose $\stmt_{12} \coarser \stmt_{34}$ and $\stmt_{14} \coarser \stmt_{23}$.
 The two pairs are incompatible (disjoint) statements whose conjunction is the certainty.
 In a sense $\stmt_{14}$ is more than half the certainty while $\stmt_{34}$ is less than half the certainty.
-So, we would expect $\stmt_{14} \broader \stmt_{34}$.
+So, we would expect $\stmt_{14} \coarser \stmt_{34}$.
 Yet, our axioms do not guarantee that.
@@
 
-One may think we are then able to create a nonsensical situation, where $\stmt_{34} \broader \stmt_{14}$.
-What happens is that now we can apply offset monotonicity multiple times like so: $\stmt_{34} \broader \stmt_{14} \implies \stmt_{3} \broader \stmt_{1} \implies \stmt_{23} \broader \stmt_{12}$.
-Therefore $\stmt_{12} \broader \stmt_{34} \broader \stmt_{14} \broader \stmt_{23} \broader \stmt_{12}$.
+One may think we are then able to create a nonsensical situation, where $\stmt_{34} \coarser \stmt_{14}$.
+What happens is that now we can apply offset monotonicity multiple times like so: $\stmt_{34} \coarser \stmt_{14} \implies \stmt_{3} \coarser \stmt_{1} \implies \stmt_{23} \coarser \stmt_{12}$.
+Therefore $\stmt_{12} \coarser \stmt_{34} \coarser \stmt_{14} \coarser \stmt_{23} \coarser \stmt_{12}$.
 Which means $\stmt_{12} \eqgran \stmt_{34} \eqgran \stmt_{14} \eqgran \stmt_{23}$.
-Therefore, while we can't derive the fact that the bigger halfs are coarser than smaller halfs, if we try to force a smaller half to be coarser than a bigger half we actually get equal halfs.
-That is: if the two halfs are comparable, then we must have at least that the bigger half is coarser than the smaller half.
-This is interesting because, in this case, the mere assumption of comparability force at least a specific relationship.
-
-So the real issue is: does it make sense for those halfs to not be comparable? It is still not clear one way or the other.
+Therefore, while we can't derive the fact that the bigger halves are coarser than smaller halves, if we try to force a smaller half to be coarser than a bigger half we actually get equal halves.
+That is, if the two halves are comparable, then we must have at least that the bigger half is coarser than the smaller half.
+This is interesting because, in this case, the mere assumption of comparability forces at least a specific relationship.
+
+So the real issue is: does it make sense for those halves to not be comparable? It is still not clear one way or the other.
 

### 2021-04-01 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: review with Andrea

@@
 
-In this section we present our main argument, which is that the state space of quantum mechanics and the space of classical probability distributions possess a parallel structure that allows one to apply commutative propositional calculus (i.e.
+In this section we present our main argument, which is that the state space of quantum mechanics and the space of classical probability distributions possess a similar structure that allows one to apply commutative propositional calculus (i.e.
 ``classical logic'') on the first and non-commutative propositional calculus (i.e.
@@
 The central point is that many features of quantum logic stem not from quantum mechanics per-se, but from the fact that quantum mechanics is a probabilistic theory, which means a classical probabilistic theory will share the same features.
-Critically, we will see that the non-commutativity of the quantum logic lattice does not originate from non-commuting observables or superposition of quantum mechanics: it is a property of the lattice of subspaces of any vector space and, indeed, of the lattice of sugroups of any group more in general.
+We find that the main point of confusion stems from comparing quantum theory, which is inherently a probabilistic theory, to a non-probabilistic classical theory.
+This approach does not allow one to understand which differences are due to the classical/quantum nature of the theory, and which are due simply to the probabilistic/non-probabilistic nature.
+By comparing quantum theory to a probabilistic classical theory, we put ourselves in a position to fully understand which difference is due to what, and have a much clearer understanding of both.\footnote{Note that probabilistic theories are important even in classical mechanics.
+Due to the nature of experimental verification, even in classical mechanics, there are always going to be uncertainties in the preparation and measurements, and repeated measurements will always lead to statistical constructs.} Critically, we will see that the non-commutativity of the quantum logic lattice does not originate from non-commuting observables or superposition of quantum mechanics: it is a property of the lattice of subspaces of any vector space and, indeed, of the lattice of sugroups of any group more in general.
+[CITE: Introduction to Lattices and Order Theory, Davey, Priestley - 2002 CUP]
 
-Mathematically, the classical commutative lattice will be provided by the Borel algebra, the smallest $\sigma$-algebra that contains the topology; the non-classical non-commutative lattice will be provided by the set of subgroups ordered by inclusion.
-We will see that the non-commutative lattice is a strict subset of the classical commutative one, meaning that everything that can be described in quantum logic can also be described in classical logic.
+Mathematically, the commutative lattice will be provided in both classical and quantum mechanics by the Borel algebra, the smallest $\sigma$-algebra that contains the topology; the non-commutative lattice will be provided by the set of subgroups ordered by inclusion.
+We will see that the non-commutative lattice is a strict subset of the commutative one, meaning that everything that can be described in quantum logic can also be described in classical logic.
 Moreover, the commutative lattice will include propositions, like propositions on the expectation value of observables, that are of scientific interest.\footnote{In an experimental context, in fact, by ``measurement'' one typically refers to the statistical collection of many ``single takes''.
@@
 
-While topologies and $\sigma$-algebras are foundational structures in mathematics for differential geometry,[CITE: Lee, Introduction to Smooth Manifolds] measure theory and probability theory,[CITE: Grimmet/Stirzaker, Probability and Random Processes] they do not seem well-known among physicists and philosophers of science, therefore we briefly summarize their definition, use and meaning.
+%they do not seem well-known among physicists and philosophers of science, therefore we briefly summarize their definition, use and meaning.
 [CITE: Kelly, Logic of Reliable Inquiry - Carcassi/Aidala, Assumptions of Physics]
 
-We start with a set $X$ which can be thought to represent the set of all possible cases we need to distinguish: all quantum states in the quantum case; all possible probability distributions in the classical case; for a simpler example, the set of the real numbers for a continuous quantity.
-Both topologies and $\sigma$-algebras are collections of subsets of $X$.
-For example, the standard topology for the real numbers will contain all the open intervals $(a, b)$.
-The difference between a topology and a $\sigma$-algebra is what types of set operations one can take within them.
+%It is useful, at this point, to review $\sigma$-algebras, which are foundational structures in mathematics,
+
+Note that mathematical branches important in the context of physics, like measure theory [CITE: Measure theory, by Donald L.
+Cohn] and probability theory,[CITE: Grimmet/Stirzaker, Probability and Random Processes] are built not top of $\sigma$-algebras, which should be considered foundational structure.
+Let us therefore review some crucial aspects.
+
+We start with a set $X$.
+In the quantum case, this will correspond to the set of all possible quantum states.
+In the classical case, this will corrsepond to the set of all possible probability distributions over phase space.
+We should stress, however, that $\sigma$-algebras are defined independently of what the elements of $X$ actually are.
+For a simpler example, $X$  could correspond to the set of the real numbers for a continuous quantity.
+A $\sigma$-algebras is a collection of subsets of $X$, closed under countable union, countable interesection and complement with respect to $X$, meaning that performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra.
+For example, a standard $\sigma$-algebras for the real numbers will typically contain all the open intervals $(a, b)$.
 
@@
 For example, the open set $(a, b)$ will exactly correspond to the statement ``the value of the quantity is between $a$ and $b$''.
-It is therefore useful to think about both the topology and the $\sigma$-algebra as collections of statements about $X$ instead of sets.
-
-A $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, meaning that performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra.
-In terms of statements, this means that it is closed under countable conjunction, countable disjunction and complement: the $\sigma$-algebra can be thought as a Boolean algebra of statements.
+It is therefore useful to think about $\sigma$-algebras as collections of statements about $X$ instead of sets.
+As the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement: the $\sigma$-algebra can be thought as a Boolean algebra of statements.
 Note that this interpretation is exactly the one used in probability.
@@
 
-A topology, instead, is closed under arbitrary union and finite intersection only.
-In terms of the statements, it is closed under arbitrary disjunction and finite conjunction.
-A topology is therefore not a Boolean algebra, but a Heyting algebra.
-Heyting algebras play a central role in intuitionist logic (and in computer science) as they describe lattices of provable (or computable) statements in finite time.
-In physics, we can understand the topology to represent statements that are experimentally verifiable in finite time.[CITE again Kelly and Carcassi/Aidala] For example, the open interval $(a, b)$ is an open set precisely because the statement ``the value of the quantity is between $a$ and $b$'' is something we can verify experimentally.
-Note that the sets in a topology are called open sets precisely because they are a generalization of the open intervals.
-On the other hand, a set with only one real value $\{ x_0 \}$ is not an open set, it is not in the topology, because the statement ``the value of the quantity is exactly $x_0$'' is not verifiable experimentally due to infinite precision.
-
-Establishing a topology, then, means establishing what statements are assumed to be experimentally verifiable.
-Given a topology, its Borel algebra is the smallest $\sigma$-algebra that contains all sets in the topology.
-That is, it is the smallest Boolean algebra that contains all the verifiable statements associated to the topology.
-Topologies and $\sigma$-algebra are pervasive in physics, to the point that all physically interesting spaces are, directly or indirectly, topological spaces.
-And wherever we have a topology, we have a Borel algebra, which is a $\sigma$-algebra, a Boolean algebra.
-
-On complete normed spaces, such as Hilbert spaces, there is a standard way to construct the topology.
+Every topological space comes already equipped with a $\sigma$-algebra called Borel algebra.
+A topology, in fact, is also a collections of sets, called open sets, and the Borel algebra is defined to be the smallest $\sigma$-algebra that contains all the open sets.
+The topology will not be directly important for us, but only as a mean to construct the Borel algebra.
+Complete normed spaces, such as Hilbert spaces, have a canonical way to construct a topology and therefore a $\sigma$-algebra.
 Let us see first how it works in a standard three-dimensional Euclidean space.
@@
 The statement ``the position is within $\epsilon$ meters of $x$'' corresponds to the set $U = \{y \; | \; |y - x|^2 < \epsilon^2 \}$ of all the points whose distance from $x$ is less than $\epsilon$.
-These are exactly the open sets that generate the topology.
+These are exactly the open sets that generate the topology, and the Borel algebra will be the smallest $\sigma$-algebra that contains these sets.
 This works in the same way for all normed spaces.
@@
 We can define a distance $\epsilon^2 = |\psi - \phi|^2=\langle \psi - \phi , \psi - \phi \rangle$ between two states, and the statement ``the state of the system is within $\epsilon$ of $\psi$'' corresponds to the set $U = \{\phi \; | \; |\psi - \phi|^2 < \epsilon^2\}$.
-As we'll see later, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.
-We will have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
+As we'll see later, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.\footnote{The choice of the square root of the density over the density itself is really a technical choice.
+Of all the $L^p$ spaces, the only Hilbert space is $L^2$, the space of the square-integrable functions.} We will have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
 In all normed spaces, the topology is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.\footnote{In the case of Hilbert spaces and quantum mechanics, the map between topology and experimental verification does not work perfectly because absolute modulus and phase are not physically meaningful.
@@
 
-From this topology, one can construct the Borel algebra.
-The Borel algebra also contains the closed sets, which are defined to be the complement of open sets.
-Recall that, in quantum logic, the propositions corresponded to closed subspaces, which are closed sets and are therefore Borel sets.
+It is a simple result that the Borel algebra thus constructed contains all propositions of quantum logic.
+
+\begin{prop}
+	Let $H$ be a Hilbert space. Let $\Sigma$ be the Borel $\sigma$-algebra induced by its inner product. Let $L(H)$ be the lattice of closed subspaces of $H$. Then $L(H) \subset \Sigma$.
+\end{prop}
+
+Proof.
+Let $p \in L(H)$ be a closed sub-space of $H$.
+Then $p$ is a closed set, the complement of an open set, and therefore a Borel sets: $p \in \Sigma$.
+Let $v \in H$ be a vector in the Hilbert space and consider the singleton $U = \{ v \}$.
+This is not a closed sub-space and therefore $U \notin L(H)$.
+As the topology of a Hilbert space is Hausdorff, all singletons are closed and therefore $U \in \Sigma$.
+This means $L(H) \subset \Sigma$.
+This concludes the proof.
+
+In quantum logic, the propositions corresponded to closed subspaces, which are closed sets and are therefore Borel sets.
 All the propositions of quantum logic are therefore readily found in the Borel algebra of the Hilbert space.
@@
 We are not taking some obscure framework from some fringe field of mathematics: these are the standard mathematical techniques used to study these spaces and can be found in any textbook.
-What we are doing here is bringing out their full physical significance, which is too often overlooked.
+[CITE: Elements of Hilbert Spaces and Operator Theory, Harkrishan Lal Vasudeva, https://www.springer.com/gp/book/9789811030192 ] What we are doing here is bringing out their full physical significance, which is too often overlooked.
 Understanding the link between the mathematical structures and the physics they represent is crucial to clearing out possible misunderstandings.
@@
 If we take two intervals $U_x$ and $U_p$ in position and momentum, and form the statement ``the state is wholly in $U_x$ and $U_p$'', meaning that the vector components are zero outside of those regions, in the classical case we always find at least one $\rho(x,p)$ that satisfies the bounds, while in quantum mechanics, if the bounds violate the uncertainty principle, no suitable $\psi(x)$ can be found.
-The point is that the differences are in the content of the logical structure, on the content of the statements, not on the type of structure.
+The point is that the differences are in the content of the logical structure, on the particular relationships among the particular statements, not on the type of structure.
 

### 2021-04-03 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: fix commutative vs distributive

@@
 
-In this section we present our main argument, which is that the state space of quantum mechanics and the space of classical probability distributions possess a similar structure that allows one to apply commutative propositional calculus (i.e.
-``classical logic'') on the first and non-commutative propositional calculus (i.e.
+In this section we present our main argument, which is that the state space of quantum mechanics and the space of classical probability distributions possess a similar structure that allows one to apply distributive propositional calculus (i.e.
+``classical logic'') on the first and non-distributive propositional calculus (i.e.
 ``quantum logic'') on the second.
-That is, both theories will have two lattices of propositions: a commutative one that follows classical logic and a non-commutative one that follows quantum logic.
+That is, both theories will have two lattices of propositions: a distributive one that follows classical logic and a non-distributive one that follows quantum logic.
 
@@
 By comparing quantum theory to a probabilistic classical theory, we put ourselves in a position to fully understand which difference is due to what, and have a much clearer understanding of both.\footnote{Note that probabilistic theories are important even in classical mechanics.
-Due to the nature of experimental verification, even in classical mechanics, there are always going to be uncertainties in the preparation and measurements, and repeated measurements will always lead to statistical constructs.} Critically, we will see that the non-commutativity of the quantum logic lattice does not originate from non-commuting observables or superposition of quantum mechanics: it is a property of the lattice of subspaces of any vector space and, indeed, of the lattice of sugroups of any group more in general.
+Due to the nature of experimental verification, even in classical mechanics, there are always going to be uncertainties in the preparation and measurements, and repeated measurements will always lead to statistical constructs.} Critically, we will see that the non-distributivity of the quantum logic lattice does not originate from non-commuting observables or superposition of quantum mechanics: it is a property of the lattice of subspaces of any vector space and, indeed, of the lattice of sugroups of any group more in general.
 [CITE: Introduction to Lattices and Order Theory, Davey, Priestley - 2002 CUP]
 
-Mathematically, the commutative lattice will be provided in both classical and quantum mechanics by the Borel algebra, the smallest $\sigma$-algebra that contains the topology; the non-commutative lattice will be provided by the set of subgroups ordered by inclusion.
-We will see that the non-commutative lattice is a strict subset of the commutative one, meaning that everything that can be described in quantum logic can also be described in classical logic.
-Moreover, the commutative lattice will include propositions, like propositions on the expectation value of observables, that are of scientific interest.\footnote{In an experimental context, in fact, by ``measurement'' one typically refers to the statistical collection of many ``single takes''.
+Mathematically, the distributive lattice will be provided in both classical and quantum mechanics by the Borel algebra, the smallest $\sigma$-algebra that contains the topology; the non-distributive lattice will be provided by the set of subgroups ordered by inclusion.
+We will see that the non-distributive lattice is a strict subset of the distributive one, meaning that everything that can be described in quantum logic can also be described in classical logic.
+Moreover, the distributive lattice will include propositions, like propositions on the expectation value of observables, that are of scientific interest.\footnote{In an experimental context, in fact, by ``measurement'' one typically refers to the statistical collection of many ``single takes''.
 Quantum logic describes the single-take measurements.
@@
 
-To really understand the source of the non-commutativity, let us study the simplest possible example.
+To really understand the source of the non-distributivity, let us study the simplest possible example.
 Let us assume that there are only two possible classical states, $1$ and $2$.
@@
 As before, the disjunction is the span, so the resulting set is different from the union, and the union of all boosts along the x-axis and the y-axis is not a subgroup and therefore not contained in the lattice.
-Thus, this has the same non-commutative structure as quantum logic.
+Thus, this has the same non-distributive structure as quantum logic.
 

### 2021-04-07 Overviews/Informational granularity/InformationGranularity.tex
InfGran: main file renamed
InfGran: biblio fixes

--- 2021-03-30 Overviews/Informational granularity/main.tex
@@
-\documentclass[10pt, onecolumn, nofootinbib]{revtex4-1}
-\usepackage[utf8]{inputenc}
-
-\usepackage[margin=1in]{geometry}
-\usepackage{amsthm}
-\usepackage{amsmath}
-\usepackage{amssymb}
-% Remove line spaces between items of enumerate and itemize
-\usepackage{enumitem}
-\setlist{noitemsep}
+\documentclass[10pt, onecolumn, longbibliography, nofootinbib]{revtex4-2}
+
+\usepackage{assumptionsofphysics}
+
+\usepackage{tikz}
 \usepackage{hyperref}
@@
 }
-\input{basicstyle}
-\input{logic}
-\input{theorems}
 \frenchspacing
 
-\newtheorem{conj}[equation]{Conjecture}
-
 \begin{document}
 
-	%TODO: add e-mail
-	%TODO: not have page break on first page
-
 \title{Assumptions of Physics blueprint: information granularity as a common foundation for geometry, measures and probability}
-\author{Gabriele Carcassi, Mark Greenfield, Christine Aidala }
+\author{Gabriele Carcassi}
+\email{carcassi@umich.edu}
+\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\author{Mark Greenfield}
+\affiliation{Math Department, University of Michigan, Ann Arbor, MI 48109}
+\author{Christine Aidala}
 \affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
@@
 The overall goal of Assumptions of Physics is to identify a handful of physical principles from which the basic laws can be rigorously derived.
-We have already established that topologies and $\sigma$-algebras can be recovered from requirements of experimental verifiability.\cite{AoPBook,AoPOverviewVerifiability} See also (TODO: Kevin Kelly).
+We have already established that topologies and $\sigma$-algebras can be recovered from requirements of experimental verifiability.\cite{aop-book,aop-overview-verifiability} See also (TODO: Kevin Kelly).
 The next step is to understand the underpinnings of geometric and measure theoretic structures (including probability) which are layered on top of topologies and $\sigma$-algebras.
@@
 Conceptually, this represents all the statements that can be associated with an experimental verification procedure.
-The full detail of why this is the correct abstraction is discussed in detail in our previous work.\cite{AoPBook} It is a property of this type of spaces that each statement $\stmt \in \tdomain$ can be expressed as a disjunction of possible cases, of the  \textbf{possibilities} of the domain.\footnote{For example, ``the mass of the object is less than 1 Kg'' is the disjunction (i.e.
+The full detail of why this is the correct abstraction is discussed in detail in our previous work.\cite{aop-book} It is a property of this type of spaces that each statement $\stmt \in \tdomain$ can be expressed as a disjunction of possible cases, of the  \textbf{possibilities} of the domain.\footnote{For example, ``the mass of the object is less than 1 Kg'' is the disjunction (i.e.
 the logical OR) of all the statements of the form ``the mass of the object is exactly $x$ Kg'' where $0 \leq x < 1$.} Formally, each possibility is a minterm\footnote{A minterm is a disjunction in which each term appears once, either negated or not.} of the countable basis.
@@
 It also means that each element can be constructed as a disjunction of a set of possibilities.
-For the full definitions, see \cite{AoPBook}.
+For the full definitions, see \cite{aop-book}.
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{apsrev4-2}
 \bibliography{bibliography}

### 2021-04-09 Overviews/Informational granularity/InformationGranularity.tex
InfGran: biblio fix
InfGran: Mark's review

@@
 
-Note that on Page 21 de Finetti\cite{definetti} specifically says that $P(E) \geq P(E’)$ implies $E \geq E’$ but not vice-versa.
+Note that on Page 21 de Finetti\cite{definetti} specifically says that $P(E) \geq P(E')$ implies $E \geq E'$ but not vice-versa.
 And that we can have $P(A)=P(B)=0$ even if $A<B$ if $A$ is impossible and $B$ is not.
@@
 And then he says that this means that the 4 postulates give a non-Archimedean structure.
-But you can make it Archimedean by ignoring the probabilities that are “infinitesimally small”: those that multiplied by a number, however big, will never become the certainty.
+But you can make it Archimedean by ignoring the probabilities that are ``infinitesimally small'': those that multiplied by a number, however big, will never become the certainty.
 On the other hand, Villegas\cite{villegas} starts already with an ordering where the infinitesimals are equal to the empty set.
@@
 
-It would be nice to be able to prove the following either true or false.
-
-\begin{conj}\label{uniformconsequence}
+We expect the following to not be provable from the given axioms:
+
+\begin{desid}\label{uniformconsequence}
     Let $\tdomain$ be uniform.
@@
 Then there exist $\stmt_2 \in \tdomain$ such that $\stmt_2 \narrower \stmt_1$ and $\stmt \eqgran \stmt_2$.
-\end{conj}
-
-[TODO: clean-up what follows]
-
-The start of a proof....
-Let $S = \{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$. We will find a maximal element with respect to fineness (OR narrowness??).
-
-The difficulty here is that for Zorn we need to find an upper bound of an arbitrary chain which also lies in $S$.
-But such an upper bound is entirely unclear on how to obtain without using the proposition to begin with.
-Uniformity is difficult to make much use of for infinite domains (beyond that $S$ is nonempty).
-
-This set $S$ is nonempty, since any possibility narrower than $\stmt_1$ must also be finer than $\stmt$ by uniformity, and the statement $\stmt_1$ must not be the impossibility since it is strictly coarser than the statement $\stmt$.
-
-Now, given any chain (with respect to narrowness), consider its disjunction. But again this may be coarser than $\stmt$!
-
-Note: problem with chains for fineness.
-Taking the disjunction of a chain with respect to fineness can yield something much coarser (i.e.
-outside of $\{\stmt[t]\in\tdomain : \stmt[t]\narrower\stmt_1 \text{ and } \stmt[t]\finer\stmt\}$) if they do not overlap.
-
-Idea: given chains in fineness, can we construct new chains which are, at each level, equigranular, but now they are also chains in narrowness?
-
-Lemma: If $C$ is a chain in fineness and $\stmt$ is such that $\stmt[t]\finer\stmt$ for all $\stmt[t]\in C$, then there exists a set $N$ and a bijection $f:C\to N$ such that $N$ is a chain in narrowness, $\OR N \narrower \stmt$, and for all $\stmt[t]\in C$, $\stmt[t]\eqgran f(\stmt[t])$.
-Use this to move chains in fineness over to chains in narrowness and then can find maximal elements in the set $S$ above ``up to equigranular equivalence classes".
-
-Problem: this is actually much harder. Prop 6 is the case of this Lemma for a chain of length 2 and the top of the chain being equal to $\stmt$.
+\end{desid}
+However, it would be nice to understand why as it would give us insights into which additional premises are needed.
+The idea is that we can always ``move" a statement to an equigranular one to sit within any other (coarser) statement.
+This would allow us to make use of various properties of narrowness.
+Interestingly, this would not give additional insight into limiting behaviors, as equigranular sequences may have vastly different limits (and so ``moving" a sequence could drastically change the equigranularity class of the limit).
 
@@
 
-\begin{conj}
-	Equivalence at $\stmt[u]$ scale is an equivalence class.
-\end{conj}
+When the definitions and axioms are solidified, the following will be essential:
+
+\begin{desid}
+	Equivalence at $\stmt[u]$ scale is an equivalence relation.
+\end{desid}
 
@@
 This would be like saying that, yes, in principles the assertions correspond to different granularity levels, but we are not to distinguish between them because the different is effectively zero.
-Understanding which options is best requires more work.
+Understanding which options are best requires more work.
 \end{remark}
 
-\begin{conj}
-	Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$, then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
-\end{conj}
+Another essential property of equivalence at $\stmt[u]$ scale is that it collapses the fineness ordering, formalized by the following statement:
+
+\begin{desid}
+	Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$. Then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
+\end{desid}
 
@@
 
-[TODO: is there anything to be salvaged between here]
-
-We start with a simpler case:
-
-\begin{prop}
+A special case of the above may be a good first goal for testing appropriate axioms, or perhaps it could end up being an axiom itself:
+
+\begin{desid}
 	Let $\stmt_1,\stmt,\stmt[v]_1\in\tdomain$ such that $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$, and suppose $\stmt[v]_1$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt\equ\stmt_1$. 
-\end{prop}
-\begin{proof}[Using Prop \ref{uniformconsequence}]
-	If both of the $\finer$'s are actually equigranular, there is nothing to show.
-Next consider the case that $\stmt_1\sfiner\stmt\finer\stmt_1\OR\stmt[v]_1$.
-Then by Proposition \ref{uniformconsequence} there exists $\stmt_2$ such that $\stmt_2\narrower\stmt$ and $\stmt_2\eqgran\stmt_1$.
-But how do we put this together??
-What about the even simpler version of the lemma with $\narrower$ instead of $\finer$?
-Just use $\stmt\OR (\stmt_1\OR\stmt[v]_1 \AND \NOT\stmt)$.
-\end{proof}
-
-This seems problematic already:
-
-We have $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$.
-We need to find some $\stmt[w],\stmt[w]_1$ infinitesimal with respect to $\stmt[u]$, such that $\stmt\OR\stmt[w]\eqgran\stmt_1\OR\stmt[w]_1$.
-Since we don't know anything about the overlap of $\stmt$ and $\stmt_1$, we can't use any operations like $\stmt\AND\NOT\stmt_1$ to find something small.
-We need to know that additional statements exist in order to find $\stmt[w]$ and $\stmt[w]_1$.
-
-For example, what if every infinitesimal statement was narrower $(\narrower)$ than $\stmt$?
-In this case, we could have something like $\stmt_1\sfiner\stmt\sfiner\stmt_1\OR\stmt[x]$ for all statements (including all infinitesimals) $\stmt[x]$.
-It would then not be the case that $\stmt\equ\stmt_1$.
-
-\begin{prop}[Need something like this but not clear if correct!]
-	Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$. Then there exist statements  $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$, infinitesimal in $\stmt[u]$, such that:
-	$$
-	\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
-	$$
-\end{prop}
-
-Note that the existence of the right combination of incompatible statements is sufficient.
-However, that is stronger than we need and may be harder to prove.
-Problem: we have very few propositions about existence of statements!
-It is hard to ensure certain ones exist.
-
-Consider the following (extreme) case. Suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$, but the infinitesimals which realize these relationships are $\stmt[v]_1$ and $\stmt[v]_2$, as follows:
-$$
-\stmt_1\OR\stmt[v]_1 \eqgran \stmt_2\OR\stmt[v]_2,\ \stmt_2\OR\stmt[v]_1\eqgran\stmt_3\OR\stmt[v]_2
-$$
-Without further axioms, it may not be possible to put this together to get transitivity (i.e. to find that $\stmt_1\equ\stmt_3$).
-
-\begin{prop}
-	The relation $\equ$ is an equivalence relation.
-\end{prop}
-\begin{proof}
-	Reflexivity and symmetry are immediate since $\eqgran$ is an equivalence relation.
-For transitivity, suppose $\stmt_1\equ\stmt_2$ and $\stmt_2\equ\stmt_3$.
-Let $\stmt[v]_1,\stmt[v]_2,\stmt[w]_2,\stmt[w]_3$ be infinitesimals with respect to $\stmt[u]$, such that $\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2$ and $\stmt_2\OR\stmt[w]_2\eqgran\stmt_3\OR\stmt[w]_3$.
-Then we have:
-	$$
-	\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2 \eqgran \stmt_2\OR\stmt[v]_2\OR\stmt[w]_2 \eqgran \stmt_3\OR\stmt[w]_3\OR\stmt[v]_2.
-	$$
-	Since $\stmt[v]_1\OR\stmt[w]_2$ and $\stmt[w]_3\OR\stmt[v]_2$ are infinitesimal with respect to $\stmt[u]$ (being finite disjuncts of infinitesimals), we have $\stmt_1\equ\stmt_3$. 
-\end{proof}
-
-[TODO: and here?]
+\end{desid}
+
+In addition to defining equivalence at scale $\stmt[u]$, we also wish to have the ordering. Eventually, the goal is to use this as a qualitative probability, as in Fishburn \cite{fishburnsurvey}.
 
@@
 
+This is the final notion of ordering.
+After narrowness orders the actual content of the statements and fineness gives a very precise ordering of size, ordering by scale in $\stmt[u]$ aims to give a less universal notion of size relative to a fixed unit statement $\stmt[u]$, where we collapse by equivalence at scale.
+
+An alternative definition is as follows.
+
 \begin{defn}\label{smalleru2}
@@
 
-\begin{conj}
+We still need to show the following:
+
+\begin{desid}
 	The two definitions are equivalent.
-\end{conj}
+\end{desid}
 
@@
 
-[TODO: would remove this]
-
-{\bf Warning:} Transitivity doesn't seem to work as stated (see the proof attempt below).
-We propose the following modification: say $\stmt_1\lequ\stmt_2$ if there exist statements $\stmt_1'$ and $\stmt_2'$ with $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ such that $\stmt_1'\finer\stmt_2'$.
-Note that {\bf by totality}, it is sufficient to just say $\stmt_1\finer\stmt_2'$ (no need for $\stmt_1'$).
-
-[TODO: proof not clear]
-
 \begin{prop}
-	$\stmt_1 \leu \stmt_2$ if for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$. 
-\end{prop}
-\begin{proof}
-	Since $\impossibility$ is infinitesimal, we have $\stmt_1 = \stmt_1\OR\impossibility \sfiner \stmt_2$, and so by Definition \ref{smalleru}, we have that $\stmt_1\lequ\stmt_2$.
+	If for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$, then we have $\stmt_1 \leu \stmt_2$. 
+\end{prop}
+\begin{proof}
+	First, using $\stmt[v] = \impossibility$ (an infinitesimal) in the hypothesis of the statement, we have $\stmt_1 = \stmt_1\OR\impossibility \sfiner \stmt_2$, and so by Definition \ref{smalleru}, we have that $\stmt_1\lequ\stmt_2$.
 Now, we must show it is not the case that $\stmt_1\equ \stmt_2$.
 
-	Let $\stmt[v]_1,\stmt[v]_2\in\tdomain$ be infinitesimals.
-Then $\stmt_1\OR\stmt[v]_1 \sfiner \stmt_2 \finer\stmt_2\OR\stmt[v]_2$.
-We conclude that it cannot be the case that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$ for any infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_1,\stmt[v]_2$.
-\end{proof}
-
-We next show that $\lequ$ is a pre-order, and that it is a partial order with respect to $\equ$-equivalence classes.
+	Suppose that $\stmt_1 \equ \stmt_2$.
+By Definition \ref{smalleru}, let $\stmt[v]_1,\stmt[v]_2\in\tdomain$ be infinitesimals such that $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$.
+But by assumption, $\stmt_1\OR\stmt[v]_1 \sfiner \stmt_2 \finer\stmt_2\OR\stmt[v]_2$.
+This contradicts the earlier equigranularity, and so it cannot be the case that $\stmt_1\equ\stmt_2$.
+\end{proof}
+
+We need it to be the case that $\lequ$ is a pre-order, and that it is a partial order with respect to $\equ$-equivalence classes.
 
@@
 
-[TODO: not clear that this works... don't we have to find infinitesimals that do not overlap?]
-\begin{proof}
-	Reflexivity is immediate since for any $\stmt\in\tdomain$, we have $\stmt\finer\stmt$, so $\stmt\lequ\stmt$. (This works for both the new and old definition)
-
-	Transitivity requires a little more work. Let $\stmt_1,\stmt_2,\stmt_3\in\tdomain$ with $\stmt_1\lequ\stmt_2\lequ\stmt_3$.
-
-	{\bf Using the original definition:}
-	We have several cases to consider.
-
-	\emph{Case 1:} $\stmt_1\finer\stmt_2\finer\stmt_3$. Then $\stmt_1\finer\stmt_3$ and so $\stmt_1\lequ\stmt_3$.
-
-	\emph{Case 2:} $\stmt_1\equ\stmt_2\equ\stmt_3$. Then $\stmt_1\equ\stmt_3$ (since $\equ$ is an equivalence relation), and so $\stmt_1\lequ\stmt_3$.
-
-	\emph{Case 3:} $\stmt_1\equ\stmt_2\finer\stmt_3$. Then for some infinitesimals $\stmt[v]_1$ and $\stmt[v]_2$ (with respect to $\stmt[u]$), we have:
-	$$
-	\stmt_1\finer\stmt_1\OR\stmt[v]_1\eqgran\stmt_2\OR\stmt[v]_2\finer\stmt_3\OR\stmt[v]_2
-	$$
-	Unclear what to do here -- need $\stmt_3$ alone on the right-hand side somehow. Case 4 has the same problem. Let's try it with the modified definition (in the ``warning").
-
-	{\bf Using the new definition:}
-	Let $\stmt_1'\equ\stmt_1$ and $\stmt_2'\equ\stmt_2$ with $\stmt_1'\finer\stmt_2'$, and let $\stmt_2''\equ\stmt_2$ and $\stmt_3''\equ\stmt_3$ with $\stmt_2''\finer\stmt_3''$. 
-	Then for an appropriate choice of infinitesimals (with respect to $\stmt[u]$) $\stmt[v]_i,\stmt[w]_i,\stmt[v]_i',\stmt[w]_i'$, $i=1,2,3$, we have:
-	$$
-	\stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_1'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \finer \stmt_2'\OR\stmt[v]_1'\OR\stmt[w]_2'\OR\stmt[w]_3'' \eqgran \stmt_2''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \finer
-	$$
-	$$
-	\stmt_3''\OR\stmt[w]_2''\OR\stmt[v]_1'\OR\stmt[w]_3'' \eqgran
-	\stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1'
-	$$
-	Since $\stmt_1 \equ \stmt_1\OR\stmt[v]_1\OR\stmt[w]_2'\OR\stmt[w]_3''$ and $\stmt_3 \equ \stmt_3\OR\stmt[w]_3 \OR\stmt[w]_2''\OR\stmt[v]_1' $, we have that $\stmt_1\lequ\stmt_3$.
-
-\end{proof}
+Fortunately, the current definitions of $\lequ$ and $\equ$ behave well together:
 
@@
 
-\begin{remark}
-	content...
-\end{remark}
-
 Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotone?
@@
 \subsubsection{Collapsing Orders}
-[TODO: didn't we find a reference for this?]
-
-Searching around did not yield the theorems desired, so we will discuss the tools needed here.
-It turns out that it is not so straightforward to collapse a poset by an equivalence relation.
-Let's start with a possible definition:
-
-If $(P,\leq)$ is a poset and $\sim$ is a collapsing equivalence relation, then endow the set $P/\sim$ with the relation $\leq/\sim$ defined by $x/\sim \leq/\sim y/\sim$ whenever there exist $x'\in x/\sim$ and $y'\in y/\sim$ with $x'\leq y'$.
-
-It turns out that this is not sufficient. Here is an important example. Let our set be $\{w,x,y,z\}$ and a poset relation defined by $x\leq w$, $y\leq w$, $y\leq z$, with the following Hasse diagram:
+In Section 2.1 of \cite{quotientposets}, the idea of collapsing a poset by an equivalence relation is introduced.
+
+If $(P,\leq)$ is a poset and $\sim$ is a collapsing equivalence relation, we may endow the set $P_{/\sim}$ with the relation $\leq_{/\sim}$ defined by $x_{/\sim} \leq_{/\sim} y_{/\sim}$ whenever there exist $x'\in x_{/\sim}$ and $y'\in y_{/\sim}$ with $x'\leq y'$ (this is the definition used in \cite{quotientposets}).
+
+It turns out that $(P_{/\sim},\leq_{/\sim})$ is not always a poset.
+Here is an important example.
+Let our set be $\{w,x,y,z\}$ and a poset relation defined by $x\leq w$, $y\leq w$, $y\leq z$, with the following Hasse diagram:
 
@@
 
-Note that neither of the above problems are possible with {\bf totality}.
-It seems that what makes the most sense is collapsing by some equivalence relation, and then extending both the ordering and the equivalence relation to generate a legitimate poset.
-We can explore this later if needed.
+Lemma 2.1.3 in \cite{quotientposets} gives sufficient conditions on the interaction of $\leq$ and $\sim$ on $P$ for $(P_{/\sim},\leq_{/\sim})$ to be a poset.
+These essentially prevent problems of the above form from happening.
+
+\begin{prop}
+Suppose $(P,\leq)$ has a minimum element $m$ and that the equivalence class $m_{/\sim}$ contains no other elements.
+Further, suppose that if $x_{/\sim} \leq_{/\sim} y_{/\sim}$, then for all $x'\in x_{/\sim}$ there exists a $y'\in y_{/\sim}$ with $x'\leq y'$.
+Under these assumptions, $(P_{/\sim},\leq_{/\sim})$ is also a poset.
+\end{prop}
+
+As for a minimum element, we have the impossibility $\impossibility$ which will be unique in its equigranularity class, but being infinitesimal it will share its $\equ$ equivalence class, so perhaps modification of the above lemma will be necessary.
 
@@
 
-\bibliographystyle{apsrev4-2}
 \bibliography{bibliography}

### 2021-04-17 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
Updates to fundamentality and density section 5
Added in a footnote on aboutness to address Gordon's worry.

Reordered some material in section 5 to be more cogent.

@@
 However, he also shows that if we consider a more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are \textit{not} categorically equivalent \parencites*[1182-83]{Barrett2}.
-Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation.} Whereas North's argument relies on a problematic criterion for comparing structure (based on the size of symmetry groups) \parencites[]{Swanson}, our argument avoids this pitfall.
-We show that a system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian.
+Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation.} Whereas North's argument relies on a problematic criterion for comparing structure based on the size of symmetry groups \parencites[]{Swanson}, our argument avoids this pitfall.
+We show that a classical system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian.
 To be Lagrangian, systems must satisfy an additional physical condition, making Lagrangian mechanics a proper subset of Hamiltonian mechanics.
@@
 The construction at the heart of our argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density.
-We show that the invariant density sits more naturally within the Hamiltonian framework, providing another reason to privilege it over Lagrangian mechanics.
+We show that the invariant density fits more naturally within the Hamiltonian framework, providing another reason to privilege it over Lagrangian mechanics.
 % Our formalism also provides a novel interpretation of the principle of stationary action within Hamiltonian mechanics.
@@
 In this case, the Lagrangian framework is less general because it requires that an additional condition hold, compared to the Hamiltonian framework.
-Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.
+Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.\footnote{One might worry that our criterion for fundamentality is subject to elementary counterexamples.
+For instance, can't we make any theory more general by adjoining arbitrary disjunctions?
+Importantly, we require that the theory remains wholly \textit{about} classical mechanics.
+Hamiltonian mechanics is more general than Lagrangian mechanics while still being about classical systems, where ``aboutness'' can be understood using Yablo's framework \parencites*[]{Yablo}.
+Having an account of \textit{aboutness} or \textit{subject matter} is a general philosophical problem that any account of fundamentality or essence must address.
+Hence, it is not a special problem facing our criterion.}
+% % What we really want is a brief formalization of ``proper subset," e.g. using partitions of possible worlds semantics. See if I can find Lewis discussion of this (maybe mentioned by Plebani)
+% % There is perhaps a further difficulty about the identity criteria for counting conditions.
+So we really just need it to be the case that all of the Lagrangian systems are a subject matter subset of the Hamiltonian systems.
+E.g.
+in the sense that the 1800s is a subset of all of history.
+need a mereological relationship.
+% Perhaps could say: this notion of a proper subset can be spelled out using Lewis on subject matter.
+Butterfield/gomes paper.
+But I could not find this discussion in their paper (might have a false memory)
+% % we could formalize this notion of ``proper subset" using what Lewis does regarding the mereology of partitions of possible worlds. This gives an account of subject matters, similar to Yablo's.
 
@@
 
-% % done; slip in a citation to Dorr/Hawthorne 2014, to note the independent metaphysical challenges that the notion of naturalness faces, and likewise for a metaphysical notion of fundamentality.
-
 % North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory.
@@
 As indicated at the start of Section~\ref{privileging}, we use ``fundamentality" in a non-metaphysical fashion.
-On our usage, differences in fundamentality amount to differences in generality.\footnote{We also embrace other non-metaphysical uses of ``fundamental," tracking differences in epistemic or conceptual primacy.
-For instance, we can think of the canonical position and momentum observables as ``epistemically fundamental" because they suffice to completely specify a classical state, fixing its values for all other observables.
-\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." Section~\ref{fundamental} discusses further senses in which the Hamiltonian formalism is more fundamental than Lagrangian mechanics.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
+Differences in fundamentality are primarily differences in generality: more fundamental frameworks require fewer conditions to hold.\footnote{We also embrace related non-metaphysical uses of ``fundamental," discussed in Section~\ref{fundamental}.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
 Compared with North's \parencites*[]{North} account of structure, our definition has two advantages.
@@
 The assumption of motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.
+Section~\ref{density} provides a further argument for viewing Hamiltonian mechanics as more fundamental.
+We argue that (i) invariant phase space densities are a fundamental physical quantity and (ii) Hamiltonian mechanics provides the proper setting for representing these densities.
+
+% % It is also a problem facing the metaphysical accounts of fundamentality, since the notion of ``essence" is also hyperintensional: e.g.
+Socrates is essentially human vs.
+Socrates is essentially (human and 2+2 = 4) [example from Stanford page].
+So we believe that we are in no worse shape than our more metaphysical opponents.
 
@@
 
+The phase space density $\rho$ is physically fundamental because it tracks how much of the system is in each possible state.
+It quantifies the amount of stuff in each phase space state, in virtue of representing the distribution.
+Typically, the canonical position and momentum observables play this role, since they suffice to completely specify a classical state; they fix the values of all other observables.
+\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." However, we will show in Section~\ref{gauge_momentum} that the conjugate momentum is gauge-redundant.
+For this reason, the invariant density provides a more fundamental specification of the state than the state variables themselves.\footnote{Section~\ref{infinitesimal} provides another reason for viewing the density as fundamental: by defining the density $\rho_{c, \xi}$, we recover the right mathematical structure, namely a symplectic manifold.
+If we did not define the density, then we would need another account of why this mathematical structure arises for classical systems.}
+
 In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
@@
 Section~\ref{dependence} shows why point-wise comparison does not hold in general: the density is a coordinate-dependent quantity, and thus it is not generically coordinate-invariant.
-Consequently, within the Lagrangian formalism, point-wise comparison holds only when we have made a special choice of coordinates.
+Within the Lagrangian formalism, point-wise comparison holds only when we have made a special choice of coordinates.
 This gives a further reason to privilege Hamiltonian mechanics.
-Section~\ref{fundamental} argues for the fundamental physical significance of the phase space density.
-Since the density is coordinate-dependent, this shows that coordinate-independence is not a necessary condition for fundamental physical significance.
+Section~\ref{fundamental} argues that since the density is coordinate-dependent, this shows that coordinate-independence is not a necessary condition for fundamental physical significance.
 Our position thereby undercuts metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms.\footnote{For arguments in favor of representing physical systems and explanations in exclusively intrinsic terms, see \textcites[]{Field,North,Chen2018}.} Section~\ref{gauge_momentum} shows why the conjugate momentum $p$ is itself a gauge-dependent quantity.
@@
 %\footnote{This illustrates how one can be mislead by the syntactical form of the mathematics to draw an interpretive conclusion that is not borne out by the physics.}
+
+% % perhaps put at end of section 5.1: We have just shown that Hamiltonian mechanics provides the proper setting for representing the density.
+By representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
 
@@
 The result is that we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in (in the sense of space-time coordinates).
-In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
+Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
+See Section~\ref{gauge} for discussion.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
 % % Try to precisely gloss/specify the condition for a coordinate system to be special in this way, such that it supports point--wise comparison.
@@
 It's a physical requirement that we must be able to express the density in some unit.
+[Sort of transcendental/regulative assumption]
 * Invariant densities: physical significance connected to coordinate invariance. [incorporated]
@@
 
-% Goal of the following paragraph is to explain the various senses in which the invariant phase space density is fundamental
-
-Section~\ref{fundamentality} characterized fundamentality in terms of generality.
-We argued that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in part because it is more general.
-By considering how Hamiltonian mechanics supports point-wise density comparisons, we have encountered two further notions of fundamentality.
-The first is epistemic: by representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
-Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
-See Section~\ref{gauge} for discussion.} Not only is this formal property methodologically advantageous, but it also bespeaks a further physical significance.
-This brings us to a second, physical notion of fundamentality: the phase space density $\rho$ is physically fundamental because it tracks how much of the system is in each possible state.
-It quantifies the amount of stuff in each phase space state, in virtue of representing the distribution.
-
-Section~\ref{infinitesimal} demonstrated a third reason for viewing the density as fundamental: by defining the density $\rho_{c, \xi}$, we recover the right mathematical structure, namely a symplectic manifold.
-If we did not define the density, then we would need another account of why this mathematical structure arises for classical systems.
-Collectively, these considerations support interpreting the invariant phase space density as not only physically significant but also fundamental.
-%see 11:50am of 12-21 meeting video for clarification of this point. There are probably finer details for me to incorporate.
-
+As argued at the start of Section~\ref{density}, the invariant density is physically fundamental because it specifies the state of the classical system.
 Yet, as Section~\ref{dependence} explained, the density is always coordinate-dependent, even when coordinate-invariant.
@@
 Affording the density fundamental significance therefore places us at odds with those who take coordinate-independence as a necessary condition for fundamentality.
+
+%We have just shown that Hamiltonian mechanics provides the proper setting for representing the density.
+By representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
+Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
+See Section~\ref{gauge} for discussion.}
+%Collectively, these considerations support interpreting the invariant phase space density as not only physically significant but also fundamental. 
+%see 11:50am of 12-21 meeting video for clarification of this point. There are probably finer details for me to incorporate.
 

### 2021-04-17 Overviews/Informational granularity/InformationGranularity.tex

@@
 A finite element has no countable infinitesimal.
-For example, take the unit interval over the reals and consider the standard Lebesque measure over the Borel sets (which would correspond to our theoretical domain).
+For example, take the unit interval over the reals and consider the standard Lebesgue measure over the Borel sets (which would correspond to our theoretical domain).
 All sets of measure zero are uncountable infinitesimals with respect to the unit: since the measure is countably additive, any disjoint union of countably many measure zero sets is going to be measure zero, therefore it will not be as big as the unit.
@@
 
+\begin{prop}
+	Let $\tdomain$ be a theoretical domain.
+If all possibilities are equigranular, then the fineness class of statements compatible with only finitely many possibilities is determined by the number of possibilities with which they are compatible.
+\end{prop}
+\begin{proof}
+	Assume all possibilities are equigranular and let $\stmt_1$ and $\stmt_2$ be compatible with finitely many possibilities.
+We show that if $\stmt_1$ is compatible with at least as many possibilities as $\stmt_2$, then $\stmt_2\finer\stmt_1$.
+We proceed by induction on $N$, the number of possibilities compatible with $\stmt_1$.
+The case of $N=1$ follows from equigranularity of possibilities.
+Now suppose the claim holds for some $N$.
+Let $\stmt_1 = p_1\OR\cdots\OR p_{N+1}$ and $\stmt_2 = q_1\OR\cdots\OR q_m$ be statements compatible with only finitely many possibilities $p_i,i=1,\ldots,N+1$ and $q_j,j=1,\ldots,m$ respectively, where $m\leq N+1$.
+
+	Consider first the case where $m\leq N$.
+Then $\stmt_2 \finer p_1\OR\cdots\OR p_N \sfiner p_1\OR\cdots\OR p_N\OR p_{N+1} = \stmt_1$, so $\stmt_2\finer\stmt_1$.
+Next, suppose $m=N+1$.
+Write $\stmt_1' = p_1\OR\cdots\OR p_N$ and $\stmt_2' = q_1\OR\cdots\OR q_N$.
+We have already that all disjunctions of $N$ possibilties are equigranular, so $\stmt_1'\eqgran\stmt_2'$.
+Possibly by rearranging the order of the possibilties in the disjunction, we may assume $q_{N+1}\ncomp\stmt_1'$ and $q_{N+1}\ncomp\stmt_2'$.
+By Proposition \ref{fineness_properties}, we have $\stmt_1\OR q_{N+1}\eqgran\stmt_2$.
+Similarly, since $p_{N+1}\eqgran q_{N+1}$, we have $\stmt_1\eqgran\stmt_1'\OR q_{N+1}$.
+We conclude that $\stmt_1\eqgran\stmt_2$.
+\end{proof}
+
+\begin{coro}
+	Let $\tdomain$ be such that all possibilities are equigranular.
+Let $x \in \tdomain$ be a possibility.
+Then there exists a unique measure $\mu_{x} : \tdomain \to \mathbb{R} \cup \{+\infty\}$ which corresponds to the counting measure of possibilities compatible to the statement.
+\end{coro}
+
 We expect the following to not be provable from the given axioms:
@@
 
-\section{Notes}
-
-\subsection{Mark Notes again again}
-
-Question: Fix $\stmt[u]\in\tdomain$, and assume the following holds.
-Let $\stmt\in\tdomain$ such that $\stmt$ is compatible with infinitely many possibilities.
-Then there exist statements $\stmt_1,\stmt_2$ with $\stmt_1\ncomp\stmt_2$ and $\stmt_1\eqgran\stmt_2$ such that $\stmt_1\OR\stmt_2\narrower\stmt$ and $\stmt\AND\NOT(\stmt_1\OR\stmt_2)$ is an infinitesimal with respect to $\stmt[u]$.
-This means we can split statements in half (up to infinitesimals, with respect to fineness).
-Further, assume $\stmt[u]$ does not admit countable infinitesimals. That is, for any countable collection $S\subseteq\tdomain$ of infinitesimals, $\OR S$ is still an infinitesimal.
-
-We wish to know whether $\leq_{\stmt[u]}$ will have monotone continuity up to infinitesimals (this may or may not be related to the above).
-While we cannot quite answer that yet since we have not proven that $\lequ$ is transitive, we can think about it for $\finer$.
-The statement to understand is:
-
-Let $\{\stmt_i:i\in\mathbb{N}\}$ be a sequence of statements monotone increasing in narrowness (i.e.
-$\stmt_i\narrower\stmt_j$ for all $i\leq j$), and write $\stmt=\OR_i\stmt_i$ be the disjunction.
-Assume that $\stmt$ is finitely comparable to $\stmt[u]$.
+\subsection{Monotone continuity up to infinitesimals}
+
+The established results in the literature of comparable probability require a specific axiom, called monotone continuity, that essentially make sure that sequences converge.
+This is necessary because their goal is to recover a probability measure that is countably additive.
+
+\begin{defn}[Monotone continuity]
+	Given a domain $\tdomain$ and a unit $\stmt[u] \in\tdomain$, we say that $\lequ$ satisfies \textbf{monotone continuity} if the following is true: given a sequence $\{\stmt_i\}_{i=1}^{\infty}$ such that $\stmt_i \narrower \stmt_j$ for all $ i \leq j$ and $\stmt = \bigOR_i\stmt_i$ the disjunction, for every $\stmt[t] \in \tdomain$ such that $\stmt_i\leq_{\stmt[u]} \stmt[t]$ for all $i$ we have $\stmt\leq_{\stmt[u]} \stmt[t]$.
+\end{defn}
+
+Note that, in general, monotone continuity is not satisfied.
+See examples in \ref{lack_of_monotone_continuity}.
+In terms of measures, the issue is that we may have that the disjoint disjuntion of countably many infinitesimals is not necessarily an infinitesimal, which means countably meany sets of measure zero would not sum up to a set of measure zero.
+
+We therefore want to make the following distinction. TODO: define countable vs uncountable
+
+\begin{defn}
+	We say that $\stmt$ is a \textbf{uncountable} infinitesimal with respect to $\stmt[u]$ if all there exists a sequences $\{ \stmt_i \}_{i=1}^\infty$ such that:
+	\begin{enumerate}
+		\item $\stmt \finer \stmt_i$ for all $i$
+		\item $\stmt_i \ncomp \stmt_j$ for all $i \neq j$
+		\item $\bigOR\limits_{i=1}^n \stmt_i \finer \stmt[u]$ 
+	\end{enumerate}
+	In words, $\stmt$ is a countable infinitesimal with respect to $\stmt[u]$ if we can find for any number, there is a disjunction of that many broader mutually incompatible statements $\stmt_i$ is finer than $\stmt[u]$.
+\end{defn}
+
+TODO: desiderata that if there are no countable infinitesimal, then we have monotone continuity
+
+We wish to understand the following. This is monotone continuity up to infinitesimals, and so with the right definitions, it should translate to monotone continuity for $\lequ$.
+
+Let $\stmt[u]$ be a statement which does not admit countable infinitesimals, and let $\{\stmt_i:i\in\mathbb{N}\}$ be a sequence of statements monotone increasing in narrowness (i.e.
+$\stmt_i\narrower\stmt_j$ for all $i\leq j$).
+Write $\stmt=\OR_i\stmt_i$ for the disjunction and assume $\stmt$ is finitely comparable to $\stmt[u]$.
 If $\stmt[t]\in\tdomain$ such that $\stmt[t]$ is finitely comparable to $\stmt[u]$, and $\stmt_i\finer\stmt[t]$ for all $i$, then $\stmt\finer\stmt[t]\OR\stmt[w]$ for some infinitesimal (with respect to $\stmt[u]$) $\stmt[w]$.
 
-Idea: added terms approach infinitesimal.
-
-Consider the new sequence defined by $\stmt'_i := \stmt_i \AND \NOT\stmt_{i-1}$ for $i\geq2$ and $\stmt'_1 := \stmt_1$.
-Then $\OR_{i=1}^n\stmt'_i = \stmt_n\finer\stmt[t]$ for all $n$.
-It follows that $\stmt = \OR_{i=1}^{\infty}\stmt'_i$, and so $\stmt$ is a countably infinite disjunction of disjoint statements.
-Since $\stmt$ is finitely comparable to $\stmt[u]$, not all of the $\stmt_i'$s can be infinitesimals.
-
-Consider $\stmt[v]_n = \OR_{i=n}^{\infty}\stmt'_i = \stmt \AND\NOT\stmt_{n-1}$.
-Observe that $$\AND_{n=1}^{\infty}\stmt[v]_n = \AND_{n=1}^{\infty}\stmt\AND\NOT\stmt_{n-1} = \stmt\AND\NOT\OR_{n=1}^{\infty}\stmt_n = \stmt\AND\NOT\stmt \equiv \impossibility$$
-
-We wish to show that $\stmt = \stmt_{n-1}\OR\stmt[v]_n\finer \stmt[t]\OR \stmt[w]$.
-Or equivalently, assuming there exist statements $\stmt[x]_n\eqgran\stmt_n$ with $\stmt[x]_n\narrower\stmt[t]$ and $\stmt[x]_{n-1}\narrower\stmt[x]_n$, we wish to show that $\stmt[w]' := \stmt[t]\AND\NOT\stmt[x]$ is an infinitesimal, where $\stmt[x]=\OR_{i=1}^n\stmt[x]_i$.
-Note: if we have a very robust collection of possibilities and a dense enough fineness relation, we can ``move statements around" to find equigranular copies inside other statements, so the claim to show reduces to the case in this paragraph.
-If $\stmt[x]$ is not narrower than $\stmt[t]$, then there must exist some possibility in $\stmt[x]$ outside of $\stmt[t]$, and this must have shown up in some $\stmt[x]_i$.
-Is this sufficient?
-
-Suppose $\{\stmt_i\}_{i=1}^{\infty}$ is such that $\AND_{i=1}^{\infty}\stmt_i = \impossibility$?
-Not sure.
-We need more theorems about characterizing infinitesimals beyond the definition.
-We don't have much in the way of finding arbitrarily large (finite) sets of equigranular statements.
-
-Comment: Let $\stmt[u] = \mathbb{R}$.
-Then let $\stmt[t] = (0,\infty)$ and $\stmt = \mathbb{R}$.
-These are both finitely comparable to $\stmt[u]$.
-Next, let $\stmt_n = (-n,n)$.
-Then this is an increasing sequence of statements in narrowness, each of which is finer than $\stmt[t]$.
-But their disjunction is $\stmt$ which is (strictly) coarser than $\stmt[t]$, and no infinitesimal $\stmt[w]$ (with respect to $\stmt[u]$) will be enough to have $\stmt\finer\stmt[t]\OR \stmt[w]$.
-Interestingly, in this case, each $\stmt_n\AND\NOT\stmt_{n-1}$ is a countable infinitesimal in $\stmt[u]$.
-This shows us that a natural space which we might like to consider (e.g.
-$\mathbb{R}$) will fail the above conclusion; however, each set $(-n,n)$ for $n\geq1$ is a countable infinitesimal in $\mathbb{R}$, and so not admitting countable infinitesimals will be an important and strict condition.
-
-Let $\stmt[u]$ be a statement.
-Suppose there exist statements $\stmt_i$, $i=1,\ldots$ which are mutually incompatible, equigranular, and $\stmt[u] = \OR_{i=1}^{\infty}\stmt_i$.
-Then all the $\stmt_i$'s are countable infinitesimals.
-So if it is possible to split a statement into countably infinitely many incompatible equigranular statements, then it admits countable infinitesimals.
-Not admitting countable infinitesimals is a strict one, but examples are provided by the standard Borel measure on bounded intervals in $\mathbb{R}$.
-
-\subsection{Unique measures}
-
-One condition for the measure to be unique given a unit statements is that all points (i.e.
-possibilities in our framework) are equigranular.
-In fact, consider a domain with just three points not equigranular to each other, but each finer than the other.
-We can construct any measure that satisfies the requirement by picking three numbers and assigning them accordingly, since we can't determine ``how bigger'' is one point with respect to the other.
-
-\subsection{Finding a probability measure}
-
-Our first goal is to find a measure which is compatible with the fineness preorder.
-The idea is that fineness tells us which sets are ``larger" (less precise) than each other, regardless of any overlap.
-A measure will quantify this and allow us to assign numbers to the sizes.
-
-The paper \cite{probexistence} seems to be extremely relevant here.
-The basic idea in this paper is to find necessary and sufficient conditions for the existence of a probability measure on a Boolean algebra (e.g.
-of subsets of some set) which is compatible with a total pre-order.
-There are a few points to keep in mind.
-
-First, we need a \emph{total} pre-order.
-This means for every pair of elements, one is finer than the other.
-This is certainly not the case for fineness in general: color and position, for example, cannot be compared.
-This is the goal of defining the \emph{unit} above.
-The unit serves as a baseline statement against which others can be compared for fineness.
-For example, a statement about position singles out length as the dimension to compare, or a statement about numbers of objects specifies ``count" as the dimension to compare (e.g.
-the discrete unit of number of objects).
-
-Next, we need the three properties specified in the paper, to be written up later! We are optimistic that these three properties are present in our situation.
-
-There is one additional caveat.
-The measure constructed by the main theorem in the above-mentioned paper is always a probability measure.
-This means the whole space has finite measure.
-For example, one could put the Lebesgue measure on $(0,1)$ but under a homeomorphism $(0,1)\to\mathbb{R}$ induce a finite (probability) measure on  $\mathbb{R}$.
-If we chose the full set as a unit, we are exactly looking for a probability measure.
-If we pick a unit that is not the full set, we can also "constrain" the problem to that subset, so that we still have a probability measure.
-Then we may be able to "stich together" the measures constructed on all equigranular sets assuming they cover the full space.
-
-\subsection{Applying results of the paper \cite{villegas}}
+\section{Adapting results from \cite{villegas}}
 
@@
 For example, the transition from a uniform sequence of finitely-valued random variables into a (single) uniform random variable makes use of a countable dense subset.
-
-\subsection{Example with lack of monotonicity}
-
-Consider a Euclidian two dimensional space charted by $\{x,y\}$.
-Let $U_{i,j}$ be the square in the interval $x \in (i, i+1)$ and $y \in (j, j+1)$.
-Let $V$ be the vertical (infinite) strip in the interval $x \in (0, 1)$ and $W$ the vertical strip in the interval $x \in (0, 2)$.
-Define the sequence $W_i = \{ U_{0,0}, U_{1,0}, U_{0,1}, U_{1,1}, U_{0,2}, U_{1,2}, U_{0,3}, U_{1,3}, ...
-\}$.
-We have $\bigcup\limits_0^\infty W_i = W$.
-This means $\bigcup\limits_0^n W\finer V$ for all finite $n$ but $\bigcup\limits_0^\infty W_i = W \coarser V$.
 
@@
 
+\subsection{Example with lack of monotone continuity}\label{lack_of_monotone_continuity}
+
+Consider the subset $[0,\infty)\times[0,\infty)\subseteq\mathbb{R}^2$.
+Let $U_{i,j} = [i, i+1)\times[j,j+1)$ be the unit square with bottom-left corner $(i,j)$.
+Let $V$ be the vertical (infinite) strip $V = [0,1)\times[0,\infty)$ and $W = [0,2)\times[0,\infty)$.
+Define the sequence $W_i = \{ U_{0,0}, U_{1,0}, U_{0,1}, U_{1,1}, U_{0,2}, U_{1,2}, U_{0,3}, U_{1,3}, ...
+\}$ building upwards row by row.
+We have $\bigcup\limits_0^\infty W_i = W$.
+This means $\bigcup\limits_{i=0}^n W_i\finer V$ for all finite $n$ but $\bigcup\limits_{i=0}^\infty W_i = W \coarser V$.
+
+TODO: show that this is a problem for $\lequ$
+
 \section{Insightful failures}
@@
 
+\subsection{Transport of statements}
+
+TODO: Make this an example of how measure zero is not equigranularity
+
+Consider the Borel sigma-algebra on $(0,1)$.
+Let $\stmt_1 = (1/2,1) \setminus \mathbb{Q}$, and let $\stmt = (0,1/3)$.
+The hope is that there would exist some $\stmt_2\narrower\stmt_1$ with $\stmt_2\eqgran\stmt$.
+In Lebesgue measure (and perhaps any reasonable measure of fineness), we would have $\stmt\finer\stmt_1$.
+However, if we try to ``move" $\stmt$ to an equigranular copy inside $\stmt_1$, it will have infinitely many ``holes" - if we wish to use other points of $\stmt_1$ to fill them, there are numerous countable subsets which we consider non-equigranular, and so it is unclear if there is any canonical choice of equigranularity class for such an $\stmt_2$.
+
 \subsection{Transport of sequences}
 
-[TODO: example on how moving a sequence to one that is equigranular at all finite steps does not give the same limit]
+Consider the following two sequences of subsets of $\mathbb{N}$.
+Let $\stmt_n = \{1,2,3,\ldots,n\}$, and let $\stmt[t]_n = \{2,4,6,\ldots,2n\}$.
+If we are treating $\mathbb{N}$ as a uniform domain, we should have $\stmt_n\eqgran\stmt[t]_n$ for all $n$.
+However, the limit $\bigOR_n\stmt_n = \mathbb{N}$, while $\bigOR_n\stmt[t]_n = \{2,4,6,\ldots\} = 2\mathbb{N} \subset \mathbb{N}$.
+These are not equigranular limits despite the sequences being equigranular.
+
+Here is another example which does not rely on countable infinitesimals.
+In the domain $[0,1]$, considered with its Borel sigma-algebra, there are no countable infinitesimals.
+Consider the sequences $\stmt_n = (1/n,1]$ and $\stmt[t]_n = (1/2n, 1 - 1/2n]$.
+These should be equigranular as they are simply offset by a shift of $1/2n$ from each other.
+Then the limits are $\bigOR_n\stmt_n = (0,1]$ and $\bigOR_n\stmt[t]_n = (0,1)$.
+While they have the same Lebesgue measure, the former has the point 1, and so it should not be equigranular to the latter.
 
 \subsection{Maximal chains}
-[TODO: boil this down to the essentials
 
@@
 This does not work.
-
-The issue is that the granularity of a countable sequence ``jumps'' at infinity.
-For example, take the sequence $\{1\}$, $\{1, 3\}$, $\{1, 3, 5\}$, ...
-, \{ all odd natural numbers\}.
-Compare to $\{1\}$, $\{1, 2\}$, $\{1, 2, 3\}$, ...
-, \{ all natural numbers\}.
-They all are maximal chains.
-They both go through all finite sets, but they reach infinity at two different levels.
-Effectively, the above property is only true over finite sets.
 
@@
 
-\begin{prop}
-Suppose there is a (partial) pre-ordering on the countably infinite proper subsets of $\mathbb{N}$ which satisfies offset monotonicity.
-Then there exists a uniform theoretical domain which lacks the maximal chains property.
-\end{prop}
-\begin{proof}
-Consider the following example.
-Let the set of possibilities be given by $\mathbb{N}$ and let the theoretical domain be the powerset $\mathcal{P}(\mathbb{N})$.
-Let fineness be defined as follows.
-All finite subsets are ordered by their cardinality (with $\impossibility = \emptyset$).
-Put any ordering on the countably infinite strict subsets of $\mathbb{N}$, and impose that they are strictly coarser than all finite sets and strictly finer than the entire set $\mathbb{N} = \certainty$.
-This is an example of a uniform theoretical domain, where all possibilities (natural numbers) are equigranular (singleton sets of cardinality 1), and the ordering of equigranular equivalence classes is total.
-
-Consider the following chain $C$ in $\tdomain$: $c_0 = \emptyset$, $c_n = \{1,\ldots, n\}$ for $n<\infty$ and write $c_{\infty} = \mathbb{N}$.
-Then $c_m \snarrower c_n$ for all $m<n$.
-We claim that this is a maximal chain (see below for an explanation).
-
-Now, notice that there are no infinite proper subsets of $\mathbb{N}$ which are in the chain $C$.
-This represents an equigranularity class which is not represented in the maximal chain $C$, which is a failure of the maximal chains condition.
-
-To see why $C$ is maximal, let $c\in\mathcal{P}(\mathbb{N})$ but that $c\neq c_{\lambda}$ for $\lambda\in\mathbb{N}\cup\{0,\infty\}$.
-Then there exists some $l\in\mathbb{N}$ with $l+1\in c$ but $l\notin c$.
-Thus $c$ is not broader than $c_p$ for any $p\geq l$, and $c$ is not narrower than $c_q$ for any $q\leq l+1$.
-It follows that for $c$ to fit in the chain, it must be higher up than $c_{l+1}$ but lower down that $c_l$, which is not possible.
-\end{proof}
+The issue is that the granularity of a countable sequence with each term a finite size but an infinite limit ``jumps'' at infinity.
+For example, take the sequence $\{1\}$, $\{1, 3\}$, $\{1, 3, 5\}$, ...
+, \{ all odd natural numbers\}.
+Compare to $\{1\}$, $\{1, 2\}$, $\{1, 2, 3\}$, ...
+, \{ all natural numbers\}.
+They are both maximal chains.
+They both go through all sizes of finite sets, but they reach infinity at two different levels.
+Effectively, the above property is only true over finite sets.
+
+The following consequences of the MCP are interesting, but only use much weaker statements about the fineness classes of possibilities.
 
 \begin{prop}\label{chainimpliesuniform}
-Let $\tdomain$ be a theoretical domain with the maximal chains property. Then the following are equivalent:
-\begin{enumerate}
-    \item the possibilities are comparable
-    \item the possibilities are equigranular
-    \item all statements compatible with only finitely many possibilities are comparable to each other.
-\end{enumerate}
-\end{prop}
-\begin{proof}
-
+Let $\tdomain$ be a theoretical domain with the maximal chains property, and suppose all possibilities are comparable. Then they must be equigranular.
+\end{prop}
+\begin{proof}
 Suppose all possibilities are comparable.
 Let $p,q$ be possibilities.
-Consider the two chains $\{\impossibility, p,\certainty\}$ and $\{\impossibility, q, \certainty\}$.
+Consider the two chains $\{\impossibility, p\}$ and $\{\impossibility, q\}$.
 By the MCP, these may be extended to maximal chains in narrowness, with corresponding maximal chains in fineness equivalence classes.
 Then since possibilities are successors of the impossibility, the fineness equivalence classes represent by $p$ and $q$ must both be an immediate successor of the equivalence class representing the impossibility.
-Now, since all possibilities are comparable, and the fineness class of $p$ is an immediate successor to that of the impossibility, we must have $p\finer q$.
-Similarly, we must have $q\finer p$.
-It follows that $p\eqgran q$.
-
-Assume all possibilities are equigranular.
-We show that any two statements compatible with only finitely many possibilities are comparable.
-In fact, we will prove that the equigranularity equivalence classes for finite disjunctions of possibilities are given by the size of the disjunction.
-We proceed by induction on $N$, the larger number of compatible possibilities for the two statements.
-The case of $N=1$ follows from equigranularity of possibilities.
-Now suppose the claim holds for some $N$.
-Let $\stmt_1 = p_1\OR\cdots\OR p_{N+1}$ and $\stmt_2 = q_1\OR\cdots\OR q_m$ be statements compatible with only finitely many possibilities $p_i,i=1,\ldots,N+1$ and $q_j,j=1,\ldots,m$ respectively, where $m\leq N+1$.
-
-Consider first the case where $m\leq N$.
-Then $\stmt_2 \finer p_1\OR\cdots\OR p_N \sfiner p_1\OR\cdots\OR p_N\OR p_{N+1} = \stmt_1$, so they are comparable, and in particular all statements compatible with fewer than $N+1$ possibilities are strictly finer than $\stmt_1$.
-Next, suppose $m=N+1$.
-Write $\stmt_1' = p_1\OR\cdots\OR p_N$ and $\stmt_2' = q_1\OR\cdots\OR q_N$.
-We have already that all disjunctions of $N$ possibilties are equigranular, so $\stmt_1'\eqgran\stmt_2'$.
-Possibly by rearranging the order of the possibilties in the disjunction, we may assume $q_{N+1}\ncomp\stmt_1'$ and $q_{N+1}\ncomp\stmt_2'$.
-By Proposition \ref{fineness_properties}, we have $\stmt_1\OR q_{N+1}\eqgran\stmt_2$.
-Similarly, since $p_{N+1}\eqgran q_{N+1}$, we have $\stmt_1\eqgran\stmt_1'\OR q_{N+1}$.
-We conclude that $\stmt_1\eqgran\stmt_2$.
-This completes the proof that $2\Rightarrow3$.
-
-The claim $3\Rightarrow1$ is immediate since 1 is a special case of 3. 
-\end{proof}
+It follows that $p\finer q$, and similarly $q\finer p$, and so $p\eqgran q$.
+\end{proof}
+
+TODO: add remark that MCP -> finite domain
 

### 2021-04-17 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
New Folders for BJPS, and cut parts of Section 5
Cut 5.1.1, 5.1.2, and 5.2. Bridged section transitions and checked for no ghost label calls

--- 2021-04-17 Articles/2020-ClassicalMechanicsPhil/Source/classical-phil-article.tex
@@
 We also consider and rebut objections stemming from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
-Along the way, we consider both the gauge redundancy of the conjugate momentum (Section~\ref{fundamental}) and of the Lagrangian (Section~\ref{gauge}).
+Along the way, we consider both the gauge redundancy of the conjugate momentum (Section~\ref{gauge_momentum}) and of the Lagrangian (Section~\ref{gauge}).
 
@@
 As indicated at the start of Section~\ref{privileging}, we use ``fundamentality" in a non-metaphysical fashion.
-Differences in fundamentality are primarily differences in generality: more fundamental frameworks require fewer conditions to hold.\footnote{We also embrace related non-metaphysical uses of ``fundamental," discussed in Section~\ref{fundamental}.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
+Differences in fundamentality are primarily differences in generality: more fundamental frameworks require fewer conditions to hold.\footnote{We also embrace related non-metaphysical uses of ``fundamental," discussed in Section~\ref{density}.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
 Compared with North's \parencites*[]{North} account of structure, our definition has two advantages.
@@
 According to Teh and Tsementzis, this reformulation shows that  ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh}.}  Additionally, each of these formalisms references some ``unphysical" coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.
-We return to the physical significance of certain coordinate choices in Section~\ref{fundamental}.
+% We return to the physical significance of certain coordinate choices in Section~\ref{fundamental}.
 %The mathematical structure that you ``need" might not be unique.
@@
 The next section describes how Hamiltonian mechanics provides the natural setting for the invariant density, which characterizes how a classical object's parts are distributed.
-Section~\ref{fundamental} argues that the invariant phase space density is more fundamental than the state variables, since conjugate momentum is a gauge-dependent quantity.
+The invariant phase space density is more fundamental than the state variables, since conjugate momentum is a gauge-dependent quantity.
 In Section~\ref{Lagrangian} we provide a geometric interpretation of the principle of stationary action within Hamiltonian mechanics.
@@
 This gives a further reason to privilege Hamiltonian mechanics.
-Section~\ref{fundamental} argues that since the density is coordinate-dependent, this shows that coordinate-independence is not a necessary condition for fundamental physical significance.
-Our position thereby undercuts metaphysically-driven demands to represent physical theories in exclusively ``intrinisic" terms.\footnote{For arguments in favor of representing physical systems and explanations in exclusively intrinsic terms, see \textcites[]{Field,North,Chen2018}.} Section~\ref{gauge_momentum} shows why the conjugate momentum $p$ is itself a gauge-dependent quantity.
+Section~\ref{gauge_momentum} shows why the conjugate momentum $p$ is itself a gauge-dependent quantity.
 This bolsters our claim that the density is more fundamental than the choice of state variable pairs.
@@
 Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
-See Section~\ref{gauge} for discussion.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.
+See Section~\ref{gauge} for discussion.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.\footnote{One might wonder if it is possible to construct a coordinate-independent density, perhaps using either i) volume forms from differential geometry or ii) the canonical volume form of Riemannian geometry.
+Both of these strategies fail, although space precludes a demonstration.}
+
 % % Try to precisely gloss/specify the condition for a coordinate system to be special in this way, such that it supports point--wise comparison.
 (Seems inappropriate to mention the Jacobian, since this matters for transforming from one coordinate system to another.)
-
-In response to this deficiency of Lagrangian mechanics, one might wonder if it is possible to construct a coordinate-independent density.
-Section~\ref{independence} considers two strategies for doing so, showing why both fail.
-The first appeals to volume forms from differential geometry.
-The second appeals to the canonical volume form of Riemannian geometry, which suffices for the treatment of many Lagrangian systems.
-In each case, we will see that the coordinate-dependence of the physically significant density $\rho$ is unavoidable.
-Due to this coordinate-dependence, the density is in general not coordinate-invariant.
-Readers satisfied with the discussion so far can proceed to Section~\ref{comparing}, which illustrates one interpretive benefit of focusing on the density: it provides a better understanding of the relationship between Hamiltonian and Lagrangian mechanics within the hyper-regular domain.
-
-\subsubsection{A coordinate-independent density?}
-\label{independence}
-
-In differential topology, a different mathematical object is called the `density,' namely a volume form (i.e.
-an alternating function from $n$ vectors to a real number) that is unaffected by changes in orientation.\footnote{The differential density generalizes volume forms in the case of a non-orientable manifold.
-However, we focus only on cases where the manifold is orientable.
-To simplify the exposition, we elide subtle differences between volume forms and differential densities, concentrating on the former.} Our physically significant density $\rho$ is actually the component of a volume form $\nu = \rho(x^i)  \epsilon dx^1 \wedge dx^2 \wedge ...
-\wedge dx^n$ where $\epsilon$ is $1$ or $-1$ depending on the orientation.
-Note that $\nu$ combines the density and the volume into a coordinate-free object.
-Considering the total mass in the region $U$, we have $m(U) = \int \nu$ so essentially $\nu = dm$, i.e.
-the infinitesimal mass contribution we are integrating.
-But this means that the value returned by $\nu$ depends not only on how much stuff there is at a given point, but also on the size of the region identified by the $n$ vectors given to the form.
-In other words, $\nu$ has dimensions of only mass and is defined as a function of both a point and a volume element.
-In contrast, the density $\rho$ is just a function of the coordinates at a point, not the volume element, and has dimensions of mass over units.
-This shows that the coordinate-independent function $\nu$ cannot play the role of the coordinate-dependent density $\rho$.
-
-In Riemannian geometry, we have a canonical volume form $dV = \epsilon \sqrt{|g_{ab}|} dx^1 \wedge dx^2 \wedge ...
-\wedge dx^n$ where $|g_{ab}|$ is the determinant of the metric tensor as expressed in local coordinates.
-In component notation, we can express the volume form as $\sqrt{|g_{ab}|}\epsilon_{a_1 ...
-a_n}$, where $\epsilon_{a_1 ...
-a_n}$ is the Levi-Civita symbol.
-This gives us an objective way to assign volumes to regions of space.
-We can then express any differential volume form $\nu = \hat{\rho} dV$ as the product of the volume form times a scalar function $\hat{\rho} \equiv \rho \ \epsilon \sqrt{|g_{ab}|}$.
-At first glance, this expression misleadingly suggests that $\hat{\rho}$ is a coordinate-independent density: it is a scalar quantity with the correct dimensions of mass over volume, meeting our standard criteria for a density.
-However, note that the numerical value of $\hat{\rho} $ equals the value of $\rho$ only if $|g_{ab}| = 1$, where $1$ is a pure number.
-This match happens only in special coordinate frames and, for this reason, $\hat{\rho} $ remains coordinate-dependent.
-In other words, the metric tensor defines a unit for the volume form (e.g.
-meters-cubed), and we always express the volume in those units, regardless of what coordinates we chose.
-This would be a natural choice of volume in a Cartesian system, for example, but not in a spherical one.
-%That is: to This is only possible if the determinant of $g_{ab}$ equals one. Hence, we must define $\hat{\rho} $ in a special coordinate system, namely one where $|g_{ab}| = 1$. 
-%det g_ab =1 means cells you construct are equal volume. whereas cartesian means not just equal volume but sides (like cubes)
-%Fortunately, we can always find a set of coordinates that, at least locally, makes $|g_{ab}|$ constant.
-Since the square root of the determinant $|g_{ab}|$ is the Jacobian of the transformation to those coordinates, $|g_{ab}|$ will be constant only when this Jacobian equals one.
-
-For example, the metric tensor of a two dimensional space may define length in terms of meters.
-The canonical volume form would then define areas in meters squared.
-This would be the natural choice of volume unit in Cartesian coordinates, where in fact $|g_{ab}| = 1$ and the volume form $dV = dx \wedge dy$.
-In polar coordinates, the natural choice of units would be meter $\cdot$ radians.
-However, the volume form will still return meters-squared, since we will have $|g_{ab}| = r^2$ and $dV = r dr \wedge d\theta$.
-This means that $\hat{\rho}$ at best expresses the density in a particularly convenient Cartesian-like coordinate system: one in which the volume is expressed in special units and we can compare the density point-wise.\footnote{This is similar to what happens with the concept of proper time: it is defined in the co-moving frame, so it is coordinate-dependent.
-Yet, the expression $d\tau = \frac{1}{c} \sqrt{g_{ab} dx^a dx^b}$ is coordinate-invariant.
-This coordinate-invariance derives from the fact that we have conventionally agreed to define proper time in terms of the object's co-moving frame, rather than in terms of a different frame.} By conventionally agreeing to work in a coordinate system where $|g_{ab}| = 1$, we can make the density $\hat{\rho}$ coordinate-invariant.
-But, the property of point-wise comparison remains coordinate-dependent: it requires working in a particular coordinate system.
-Outside the context of Hamiltonian mechanics, the best we can do is obtain the property of point-wise comparison in a special coordinate system.
-% General idea from one-11 meeting: in the general case, in order to make the density a coordinate-invariant [so that we all can agree on its value], we have to make a global choice of coordinate system.
-We need to choose coordinate such that the determinant of the metric tensor is equal to 1.
-So we are seemingly saying that there are coordinate systems that are special, namely those where we can do point-wise comparison.
-So this poses a problem for the view that nature does not privilege one coordinate system over another.
-Possible response from the metaphysicians: perhaps the real fundamental physical object is a quotiented-out version of the density.
-
-% From before 1-11 : Again, the upshot here is that we cannot avoid using a coordinate-dependent density. 
-%We can also see this if we look at the units of $\hat{\rho}$, which are in terms of inverse volume as expressed in the privileged system.
-From 1-4 meeting: I thought that $\hat{\rho}$ has units of mass over volume.
-% Original bad explanation (see 1-11-21 meeting, around 11:13 am): Nevertheless, $\hat{\rho}$ remains coordinate-dependent because it is a function of $|g_{ab}|$, which depends on the choice of coordinates and is, in general, a function of the point.
-
-\subsubsection{Comparing densities on the tangent vs. cotangent bundles}
-\label{comparing}
-
-% % Wonder if it is possible to move this material to the `interpreting Lagrangian mechanics' section?
-
-Focusing on the density has the further benefit of elucidating the relationship between Hamiltonian and Lagrangian mechanics, especially within the hyper-regular domain.
-Consider a distribution over position and velocity, i.e.
-on the tangent bundle in Lagrangian mechanics.
-Note that the velocities transform in the same way (i.e.
-contravariantly) as differentials in position.
-Since both position and velocity transform contravariantly, the density will transform according to $\rho(x^i, v^j) = \sqrt{|g_{ab}|} \sqrt{|g_{ab}|} \rho(\hat{x}^i, \hat{v}^j) = |g_{ab}| \rho(\hat{x}^i, \hat{v}^j)$.
-The canonical volume form for the tangent bundle becomes: $dV_{TM} = \epsilon \sqrt{|g_{ab}|} dx^1 \wedge dx^2 \wedge ...
-\wedge dx^n \bigwedge \epsilon \sqrt{|g_{ab}|} dv^1 \wedge dv^2 \wedge ...
-\wedge dv^n = |g_{ab}| \bigwedge dx^i \bigwedge dv^j$.
-Hence, the determinant of the metric tensor equals the component of the volume form of the tangent bundle.
-The relationship to the density as expressed in phase space is therefore the following:
-\begin{equation}
-	\rho(q^i, p_j) = m |g_{ab}| \rho(x^i, v^j).
-\end{equation}
-A change of unit-variables will induce a contravariant change in the velocity and a covariant change in the momentum.
-The left side will remain invariant.
-The two terms $|g_{ab}|$ and  $\rho(x^i, v^j)$ on the right side will change in the opposite way so that their product is invariant.
-Focusing on densities, then, brings about a physically-motivated way to understand the relationship between the cotangent and the tangent bundle, where the metric tensor plays a role in defining the geometry of the latter similar to how the symplectic form defines the geometry of the former: it defines the count of states so that densities can be properly tracked.
-
-These relationships also allow us to better understand the hyper-regular case, and in particular the case of a massive particle.
-Note that, in general, the relationship between densities over the cotangent and tangent bundles is, as we saw:
-\begin{equation} \label{densities_relationship}
-\rho(q^i, p_j) = \left|\frac{\partial v^j}{\partial p_k}\right| \rho(x^i, v^j)  = \left|\frac{\partial H}{\partial p_k \partial p_j}\right| \rho(x^i, v^j).
-\end{equation}
-In the case of massive particles, where the Hamiltonian (and the Lagrangian) are quadratic, the Hamiltonian of Equation~\ref{massive_Hamiltonian} entails that $\frac{\partial H}{\partial p_k \partial p_j} = m g^{kj}$.
-This is the case where the densities in the tangent bundle can be properly understood in terms of Riemannian geometry.
-For the more general case, we note that
-\begin{equation}
-\begin{aligned}
-\delta^i_j = \frac{\partial v^i}{\partial v_j} = \frac{\partial v^i}{\partial p_k}\frac{\partial p_k}{\partial v_j} &= \frac{\partial}{\partial p_k}\left(\frac{\partial H}{\partial p_i}\right) \frac{\partial }{\partial v_j} \left( \frac{\partial L}{\partial v_k} \right) \\
-\Rightarrow \frac{\partial L}{\partial v^k \partial v^j} &= \left[ \frac{\partial H}{\partial p_k \partial p_j} \right]^{-1}
-\end{aligned}
-\end{equation}
-which means that the Hessian of the Hamiltonian is the inverse of the Hessian of the Lagrangian, and so are their determinants.
-Recall that the hyper-regular case corresponds to the case where the determinant of the Hessian of the Lagrangian is non-zero.
-Hence, Equation~\ref{densities_relationship} tells us that in this case, we can express the density of states over position and velocity.
-
-% Remember: the density is coordinate-invariant but not coordinate-independent; it is actually coordinate-dependent (i.e. it does not admit of something like a coordinate-free definition)
-% Also remember: there are quantities that are both coordinate-variant and coordinate-dependent that are still physically significant, e.g. choices of q's and p's (?)
-% Coordinate-invariant in general means that definition of the quantity involves referencing coordinates, but if you change the coordinates, then the value of the quantity remains the same.
-(E.g.
-transforms as a scalar).
-so densities are never coordinate-independent.
-defining a density requires a choice of units, so involves referencing units.
-%see 57.3 minutes of 12-21 recording: when we define the density originally, it is coordinate-free (rho as a function of the state space of the particles, S).
-But when we go to express it numerically, then it becomes coordinate-dependent.
-(I.e.
-when we express it over a manifold, then it needs to be a density for integration, and hence it becomes coordinate-dependent).
-% see 51 minutes of 12-21 recording: Densities (the coordinate-dependent kind) are physically significant partly because they tell you how much stuff is in each state
-
-%\subsection{Fundamental density and gauge-dependent momentum}
-\subsection{Coordinate-dependent yet fundamental} %; Coordinate-variant yet significant
-\label{fundamental}
-
-\begin{comment}
-
-Section on Ham. Not everything physical has to be coordinate independent or invariant.
-
-* In phase space, we do have coordinate invariance and point-wise comparison.
-Momentum changes covariantly.
-It is the density that is truly fundamental, not the coordinates in general.
-Though, the q's are priviledged from epistemic access.
-The p's are not, in fact we can define gauge transformation as things that keep the q's the same but change the p.
-
-*[Incorporate the following from section 5.2] Not all ``state variables" are physical.
-Hamiltonian mechanics does not really care what the q's and p's are.
-It is the motion assumption that tells us that q's are position, simply because they define the unit system.
-P is a gauge quantity (we care about the trajectory, the velocity).
-In this view, p is simply a ``label'' we can use to express the density over states.
-Therefore we do care about the physicality of the invariant distribution more than the physicality of the state variables themselves.
-$->$ Invariant density is really the important physical element/object in Hamiltonian mechanics.
-
-* Units are physically significant.
-In fact, the ability to express the densities in units is what gives us the symplectic structure.
-It's a physical requirement that we must be able to express the density in some unit.
-[Sort of transcendental/regulative assumption]
-* Invariant densities: physical significance connected to coordinate invariance. [incorporated]
-* Density comparison is independent of coordinates. We can compare densities point-wise without worry. [incorporated]
-
-* Makes the fundamental role that phase space plays in statistical mechanics more physically salient/perspicuous.
-Areas/volumes are measures of number of states.
-They are coordinate invariant.
-Also the Shannon/Gibbs entropy are coordinate invariant.
-
-% Section~\ref{action} uses the invariant density to provide a geometrical interpretation of the action principle.
-
-\end{comment}
-
-As argued at the start of Section~\ref{density}, the invariant density is physically fundamental because it specifies the state of the classical system.
-Yet, as Section~\ref{dependence} explained, the density is always coordinate-dependent, even when coordinate-invariant.
-In other words, there is no coordinate-independent (or coordinate-free) definition of the invariant phase space density.
-Affording the density fundamental significance therefore places us at odds with those who take coordinate-independence as a necessary condition for fundamentality.
-
-%We have just shown that Hamiltonian mechanics provides the proper setting for representing the density.
-By representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
-Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
-See Section~\ref{gauge} for discussion.}
-%Collectively, these considerations support interpreting the invariant phase space density as not only physically significant but also fundamental. 
-%see 11:50am of 12-21 meeting video for clarification of this point. There are probably finer details for me to incorporate.
-
-Proponents of this necessary condition might respond by noting that although the density $\rho_{c, \xi}$ is not coordinate-independent, the distribution $\rho_{c}$ is.
-For the distribution is equivalent to the classical state, which is a thing in the world, i.e.
-an ontic entity.
-We grant that the distribution is ontically fundamental in a sense that does not accrue to the density (which represents the distribution).
-However, we maintain that the density itself is of fundamental significance, both epistemically and physically.
-For it is only through the density $\rho_{c, \xi}$ that we gain epistemic access to the distribution.
-Although the choice of coordinates $\xi$ is fairly arbitrary, we need to make some choice in order to quantitatively measure the distribution $\rho_{c}$ through experiments.
-Additionally, defining the density is necessary for deriving symplectic structure, as shown in Section~\ref{infinitesimal}.
-Insofar as our opponent views the symplectic structure of Hamiltonian mechanics as fundamental, they must view the density as equally fundamental---despite its coordinate-dependence.
-For the same reason, insofar as our opponent refuses to grant the density fundamental status, they will have to withhold this same status from the symplectic structure.
-Hence, the coordinate-independent notion of fundamentality needs to be supplemented with a coordinate-dependent, epistemic notion of fundamentality.
-This latter notion is no less important, and it shows that it is a mistake to desire a completely ``intrinsic" formulation of a physical theory.
-% % Further claim: the ontic, coordinate-independent notion of fundamentality is insufficient for doing physics.
-We need a corresponding epistemic notion of fundamentality, since we units, state variables, and the density in order to connect the distribution to experiment.
-One can argue that any particular choice of state variables or units or coordinates for the density is arbitrary, but we need some arbitrary choice in order to quantitatively measure the distribution through experiments.
-% % Someone might also worry that the density cannot be more fundamental than the state variables, since we define the density over state variables.
-In response, we can say the following: whereas the density is coordinate-invariant, the state variables are not.
-So the state variables are more arbitrary than the density.
-Note that we do not directly measure states of particles.
-We measure quantities.
-So the only way to measure the distribution is to express it as a density in terms of some quantities, i.e.
-in terms of some state variables.
-So we need the state variables to give epistemic access to the distribution, via the density.
-So the state variables themselves are not important (e.g.
-we can choose the unit basically however we want, and the conjugate momentum has even more freedom, due to gauge redundancy).
-The physically important thing is access to the distribution, which we gain through the density.
-The classical system is made up of pieces and the density lets us say how many pieces we have in various states.
-% A defender of the necessity of coordinate-independence for fundamentality might respond as follows:
-
-% % [To do: provide literature references for this and motivating quotations; likewise for earlier remark about intrinsicality: Field, North, Chen]
-
-%[XXX To do: insert discussion of statistical mechanics and entropy here: statistical mechanics works best on phase space because phase space gives us a uniform measure over q and p.
-Focusing on the density makes the fundamental role that phase space plays in statistical mechanics more salient/perspicuous.
-Areas/volumes are measures of the number of states.
-They are coordinate invariant.
-Also the Shannon and Gibbs entropies are coordinate invariant.]\footnote{For further discussion of the importance of phase space for statistical mechanics and entropy, see \textcites{shpmp} [omit this for review?].}
-
-Instead, of coordinate-independence, we take coordinate-invariance as a weaker necessary (but not sufficient) condition for fundamental physical significance.
-Being necessary for fundamental significance does not, however, make coordinate-invariance necessary for \textit{non-fundamental} physical significance.
-Many representational choices are coordinate-variant yet still physically significant.\footnote{Indeed, many physically significant quantities are gauge-dependent, such as conjugate momentum and the Lagrangian (see Sections~\ref{gauge_momentum} and~\ref{gauge}).
-Insofar as gauge-dependency is analogous to being coordinate-variant, this supplies a further reason for denying that coordinate-invariance is a necessary condition for physical significance.} Our position here again places us at odds with others in the literature.
-For instance, North claims that coordinate choices are ``merely arbitrary choices of description" \parencites*[61]{North}.
-However, some choices of coordinates are particularly well-suited to certain problems, such as the use of action--angle variables for integrable systems, or spherical coordinates for systems with spherical symmetry.
-Similarly, we saw in Section~\ref{dependence} that some coordinate choices in Lagrangian mechanics are physically significant because they enable the property of point-wise comparison.
-This is related to the fact that, in Lagrangian mechanics, Cartesian coordinates are a physically significant coordinate choice because they decompose position and velocity variables as independent degrees of freedom.
-This independence does not hold generically across coordinate systems.\footnote{In contrast, canonical position and conjugate momentum variables always characterize separable/independent degrees of freedom in the Hamiltonian framework.} In a similar vein, the next section shows that the conjugate momentum is gauge-dependent, but it is presumably still physically significant.
-%Realized that the following remarks probably go with earlier argument about coordinate-dependence: Finally, as the derivation in Section~\ref{infinitesimal} shows, a necessary condition for defining the symplectic structure is that we can express the density $\rho_{c, \xi}$ in a particular choice of units.
-If we did not have the physical ability to define the density in some units we would not be able to define this structure.
-This provides another instance of a physically significant but coordinate-dependent representational choice.
-
-% % Could insert a summary paragraph either here or at the end of the next subsection explaining the following: coordinate-invariance and gauge-invariance are both necessary for fundamental physical status.
-But they're not necessary for being physically significant.
-Both coordinate-variant and gauge-dependent quantities can still be physically significant, although they cannot be physically fundamental.
-So in this way, we view gauge-dependency as a kind of coordinate-variance (rather than a kind of coordinate-dependence).
-And gauge-invariance is analogous to a kind of coordinate-invariance, rather than a kind of coordinate-independence.
-And perhaps this helps explain the physical significance of gauge-variant quantities.
-% Question: is any quantity that is coordinate-variant automatically coordinate-dependent?
-If so, then is coordinate-variance a special case of coordinate-dependence?
-And so gauge-variance would also be a special case of coordinate-dependence?
 
@@
 To see this, recall that the phase space density remains invariant upon changing the unit-variables $q$, inducing a state variable transformation on the conjugate momenta $p$ (which compensate by transforming covariantly).
-It is in this sense that the unit-variables $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.\footnote{As discussed in Section~\ref{fundamental}, these units are physically significant despite not being coordinate-invariant.} Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are of equal physical significance.
+It is in this sense that the unit-variables $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.
+Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are of equal physical significance.
 In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
 Unlike both velocity and the kinetic momentum $m g_{ji} v^i$, conjugate momentum is not an experimentally measurable quantity, because its specification involves a gauge choice.
+%\footnote{As discussed in Section~\ref{fundamental}, these units are physically significant despite not being coordinate-invariant.}
 
@@
 
-As discussed in Section~\ref{fundamental}, the conjugate momentum $p$ is a gauge-dependent variable.
+As discussed in Section~\ref{gauge_momentum}, the conjugate momentum $p$ is a gauge-dependent variable.
 In contrast, position and velocity are both gauge-invariant, making them physically privileged.

### 2021-04-18 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
Proofread and light edits
Checked to remove any remaining references to sections I deleted for space. Caught some typos.

@@
-\documentclass[letterpaper]{article}
+\documentclass[12pt, english, twoside]{article} %twoside necessary for fancyheader odd vs. even commands
+
+\usepackage[letterpaper]{geometry}
+%\geometry{verbose, a4paper, tmargin=1in,bmargin=1in,lmargin=1.25in,rmargin=1.25in}
+\geometry{verbose, a4paper, tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
 
 \usepackage{amsmath, hyperref, enumerate, comment}
-\usepackage{amsfonts, esint}
+\usepackage{amsfonts, esint, setspace}
 \usepackage{graphicx}
@@
 
-\usepackage[backend=biber, style=authoryear, url=false, doi=false, eprint=false]{biblatex}
-%\usepackage[backend=biber,  style=authoryear-icomp, url=false, doi=false, eprint=false]{biblatex}
+\linespread{1.3}
+
+\raggedbottom % prevents random spaces between paragraphs from appearing
+
+\usepackage[titletoc,title]{appendix}
+
+% headers
+\usepackage{fancyhdr} 
+\pagestyle{fancy}
+\fancyhead{}
+\fancyhead[OR,ER]{\large \textbf{\thepage}}
+%\fancyhead[C]{\textit{}}% odd page header and number to right top
+%\fancyhead[EC]{\textit{Your Name}}%Even page header and number at left top
+\fancyfoot[L,R,C]{}
+\renewcommand{\headrulewidth}{0pt}
+%\renewcommand{\footrulewidth}{0pt}
+\renewcommand*\footnoterule{} %removes footnote line
+
+% % redefines abstract to be capitalized, with no margins 
+\renewenvironment{abstract}
+ {\small
+  \begin{center}
+\normalsize  \textnormal{ABSTRACT\\} \vspace{-0em}\vspace{0pt}
+  \end{center}
+  \list{}{%
+    \setlength{\leftmargin}{0in}%
+    \setlength{\rightmargin}{\leftmargin}%
+  }%
+  \item\relax}
+ {\endlist}
+
+% section headings
+\usepackage{sectsty}
+\allsectionsfont{\large\raggedright\centering}
+
+% Table of contents stuff and footnotes 
+\setcounter{tocdepth}{2}
+\usepackage[]{tocloft}
+%\addtocontents{toc}{\cftpagenumbersoff{section}} % turns off page numbers from section level 
+%\addtocontents{toc}{\cftpagenumbersoff{subsection}} %like wise for subsection level 
+\renewcommand{\cftsecfont}{\itshape}
+\renewcommand{\cftsubsecfont}{\itshape}
+\renewcommand{\cftsecpresnum}{\normalfont \bfseries} % puts the section numbers in non-italics and bold
+\renewcommand{\cftsubsecpresnum}{\normalfont \bfseries} % likewise for subsection numbers
+
+\renewcommand{\cftbeforesecskip}{3pt} %controls spacing between ToC items 
+\makeatletter
+\renewcommand{\@cftmaketoctitle}{}
+
+% % footnote modification: must be between \makeatletter and \makeatother
+
+\renewcommand{\@makefntext}[1]{%
+  \setlength{\parindent}{0pt}%
+  \begin{list}{}{\setlength{\labelwidth}{6mm}% indents footnote text 
+    \setlength{\leftmargin}{\labelwidth}%
+    \setlength{\labelsep}{5pt}% moves footnote mark to the left
+     \setlength{\itemsep}{0pt}%
+      \setlength{\parsep}{0pt}%
+      \setlength{\topsep}{-3pt}% removes separation between footnotes
+    \footnotesize}%
+  \item[\@textsuperscript{\@thefnmark}\hfil ]#1% @makefnmark
+  \end{list}%
+}
+%      \setlength{\footnotesep}{0pt}%
+
+\makeatother
+\renewcommand{\cftdot}{} % removes dots from table of contents
+
+% % for fonts:
+\usepackage[T1]{fontenc}
+\usepackage[utf8]{inputenc}
+\usepackage{mathptmx}
+
+% Hyperref set-up & metadata
+\hypersetup{
+	pdftitle={Your Paper Title},
+	pdfauthor={Your Name},
+	pdfsubject={},
+	pdfkeywords={},
+	pdfcreator={XeLaTeX},
+	pdfproducer={XeLaTeX},
+	pdftoolbar=false,	% show Acrobat’s toolbar?
+	pdfmenubar=true,	% show Acrobat’s menu?
+	pdffitwindow=false,	% window fit to page when opened
+	pdfstartview={FitH},	% fits the width of the page to the window
+	unicode=true,		% non-Latin characters in Acrobat’s bookmarks
+	pdfnewwindow=true,	% links in new window
+	colorlinks=true,	% false: boxed links; true: colored links
+	linkcolor=black,	% color of internal links
+	citecolor=black,	% color of links to bibliography
+	filecolor=black,	% color of file links
+	urlcolor=blue		% color of external links
+}
+
+%\usepackage[backend=biber, style=authoryear, url=false, doi=false, eprint=false]{biblatex}
+
+% % % Biblatex commands for Citations and References:
+\usepackage[backend=biber, style=ext-authoryear-comp,  maxcitenames=2, dashed=false, url=false, doi=false, eprint=true, giveninits=true, innamebeforetitle=true, maxbibnames=6]{biblatex}
+
 \addbibresource{bibliography.bib}
 
+%\addbibresource{templateReferences.bib} %%generated by reference items above
+%maxcitenames=2 replaces all instances of more than 2 authors in citations with `et al.', but lists all authors in references
+%labeltitleyear=true; sortcites=true
+%innamebeforetitle=true places editors before the book titles
+%style=authoryear-comp adds in some weird ``In:'' stuff in reference list that must be changed below.
+
+\DeclareFieldFormat{labeldate}{\mkbibbrackets{#1}} %puts brackets around years in citations
+
+\DeclareFieldFormat{postnote}{\mkpageprefix[pagination][\mkcomprange]{#1}} % compresses page range in citations, so that unchanged digits are not repeated.
+See p.
+259, section 4.6.4 of biblatex manual.
+%also adds pagination prefix p.
+
+% clears un-needed fields from reference items
 \AtEveryBibitem{%
@@
   \clearfield{series}%
-  \clearfield{pages}%
+ \clearfield{number}%
+ \clearfield{pagetotal}%
+ \clearfield{chapter}%
   }
+  %\clearfield{pages}%}
+
+  \renewcommand{\newunitpunct}{\addcomma\space} %makes a comma the default separator between items; overridden in some cases by commands below
+
+\renewcommand{\intitlepunct}{\addspace\nopunct} %removes punctuation after ``in"
+
+\DeclareDelimFormat[bib,biblist]{nametitledelim}{\addcolon\space} %puts colon after year
+
+\DeclareFieldFormat{biblabeldate}{\mkbibbrackets{#1}} %puts brackets around years in references
+
+\DeclareFieldFormat{editortype}{\mkbibparens{\textit{#1}}} %puts ed in parentheses
+\DeclareDelimFormat{editortypedelim}{\addspace}%prevents punctuaction clash after (ed)
+\DeclareDelimFormat[bib,biblist]{innametitledelim}{\addcomma\space} % puts comma after (ed)
+
+\DeclareFieldFormat{eprint}{\printtext{available at}\addspace \textless#1\textgreater} % Calls url from Eprint field. See biblatex manual, section 4.11.2 Electronic Publishing Information, page 310-11
+
+\DeclareFieldFormat[article,periodical]{volume}{\textbf{#1}} %puts volume in bold; note that adding \bibstring{jourvol}~ before #1} prints ``vol.". from biblatex-ext manual 
+%\DeclareFieldFormat[article,periodical]{number}{#1}%\bibstring{number}~ before #1} prints ``vol."
+\renewcommand*{\jourvoldelim}{\addcomma\space} %combined with above command, puts comma after article title and before journal volume. 
+%\renewcommand*{\volnumdelim}{\addcomma\space}
+
+\DeclareFieldFormat[inbook, incollection, book]{volume}{\bibcpstring{volume}~#1} %capitalizes `Vol' after book titles.
+
+\DeclareFieldFormat{pages}{\mkpageprefix[pagination][\mkcomprange]{#1}} % compresses page range in bibliography, so that unchanged digits are not repeated.
 
@@
 
+\newcommand\bs{\begin{singlespace}} 			
+\newcommand\es{\end{singlespace}}
+
 \begin{document}
-
-\title{From physical conditions to Hamiltonian and Lagrangian mechanics}
-\author{TBD}
-
-\date{\today}
+\title{\vspace{-2em} Hamiltonian Privilege}
+%\author{TBD}
+
+\date{\vspace{-3em} \today}
 
@@
 
+% 300 words maximum for BJPS
+\vspace{-3em}
+\bs
 \begin{abstract}
-	TBD
+\noindent We argue that Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
+Our argument provides a non-metaphysical strategy for privileging one formulation of a theory over another: a formulation is more fundamental at least when it is more general.
+We illustrate this criterion through a novel interpretation of classical mechanics in terms of three physical conditions.
+Two of these conditions suffice for recovering Hamiltonian mechanics.
+A third condition is necessary for recovering Lagrangian mechanics.
+This shows that Lagrangian mechanics is a proper subset of Hamiltonian mechanics, elucidating the connection between these two formalisms.
+Using our framework, we argue that the invariant phase space density is physically fundamental.
+Since Hamiltonian mechanics provides the proper setting for this density, we gain another reason for privileging it over Lagrangian mechanics.
+Finally, we interpret Lagrangian mechanics within the Hamiltonian framework, providing a geometric interpretation of the principle of stationary action.
+We rebut arguments for privileging Lagrangian mechanics, showing where they go wrong.
 \end{abstract}
-
+\es
+
+\bs
 \tableofcontents 
+\es
 
@@
 \label{introduction}
-
-% TODO: Lagrangian assumption vs kinematic assumption. Investigate what happens when the transformation is non-linear.
 
@@
 This equivalence motivates the view that there are no physically significant differences between these two formalisms, at least within the hyper-regular domain.
-% It is interesting to think how the other formulations fit in, e.g. Hamilton--Jacobi. Is there a categorical equivalence result for this formulation in some domain?
 
@@
 
-Section~\ref{privileging} provides an informal overview of our argument for privileging Hamiltonian mechanics.
+Section~\ref{privileging} previews our argument for privileging Hamiltonian mechanics.
 We introduce three physical conditions that define classical systems while distinguishing Hamiltonian from Lagrangian mechanics.
@@
 In contrast, Lagrangian mechanics does not support these important properties of the density.
-Despite being coordinate-invariant, one cannot make the density truly coordinate independent.
-We therefore argue that some quantities are physically significant despite being coordinate-dependent.
 Finally, Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework.
@@
 We also consider and rebut objections stemming from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
-Along the way, we consider both the gauge redundancy of the conjugate momentum (Section~\ref{gauge_momentum}) and of the Lagrangian (Section~\ref{gauge}).
-
-% We define classical systems in terms of basic physical conditions that they satisfy. These conditions justify the appropriateness of certain mathematical structures.
-
-% Understanding the relationship between Hamiltonian and Lagrangian mechanics depends partly on how one defines ``classical systems." Different definitions can result in different interpretive relationships.
-For instance, restricting classical systems to the hyper-regular domain precludes certain interpretive considerations.
-To avoid verbal disputes, it is essential to clarify what falls within the scope of ``classical systems." One way to delimit classical systems is to specify a set of equations that they satisfy---such as Hamilton's equations or the Euler--Lagrange equations.
-We will instead follow a different approach by delimiting classical systems based on the physical conditions that they satisfy.
-Section~\ref{infinitesimal} introduces the condition of \textit{infinitesimal reducibility}, meaning that the state of the whole system is equivalent to specifying the states of its parts (and vice versa).
-Infinitesimal reducibility distinguishes classical from quantum systems, where the latter are irreducible (we cannot specify the whole state by specifying the states of its subsystems).
-In the widest sense, a classical system is any infinitesimally reducible system.
-%a reducible system comprises subsystems whose states jointly determine the state of the whole system.
-
-% Section~\ref{deterministic} introduces the joint condition of deterministic and reversible evolution.
-We will show that a classical system which satisfies this condition is Hamiltonian: its phase space is a symplectic manifold that evolves according to Hamiltonian evolution  (i.e.
-according to a symplectomorphism).
-Following a deterministic and reversible evolution is equivalent to being a conservative system, i.e.
-one that conserves energy.
-Hence, our discussion of Hamiltonian mechanics necessarily excludes dissipative systems, i.e.
-ones that do not conserve energy.
-Likewise, our derivation of Lagrangian mechanics in Section~\ref{motion} will also take place within the context of conservative systems.
-Consequently, we will not consider systems that satisfy inhomogenous Euler--Lagrange equations.
-We, of course, do not deny the sense in which the Lagrangian formulation applies to dissipative systems, but this is not what we will mean by ``Lagrangian mechanics" \parencites[]{Smith}{Curiel}.\footnote{Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime.
-We will view dissipative systems as Newtonian, but neither Lagrangian nor Hamiltonian.
-
-% General decomposition of conservative and non-conservative: https://scholar.harvard.edu/files/rsyuan/files/ijbc.pdf
-
-%Further motivation for focusing on non-dissipative systems: dissipative systems contain more information than contained in the Lagrangian alone, namely, the additional information contained in the Rayleigh potential.
-Or, as Curiel puts it, the systems involve a ``force $1 $-form that is not exact and that is not Lie-derived along the Lagrangian vector field solving the inhomogeneous system" [311].
-
-%Note that the photon does not belong to the hyper-regular domain. Furthermore, if you are not in the hyper-regular domain, then there is no longer a one-to-one map between position and velocity.
-
-%It seems as though North is explicitly not assuming that her argument applies only to the hyper-regular domain [page 85, especially footnote 50].
-Part of her argument involves motivating the idea that we will not in general be able to construct an equivalent Lagrangian formulation, and this involves appealing to systems outside the hyper-regular domain.
-% To-do: Figure out how North defines classical systems.
-It might be in terms of the equations of motion that hold for classical particles [87].
-North is arguing that ``symplectic structure is a simpler, more natural, more general underlying structure" that ``suffices for" this theory of classical mechanics [87].
-Seems similar to GC position.
-
-% In the introduction, should we give glosses on hyper-regular, infinitesimally reducible, deterministic/reversible, and motion equivalence? Otherwise, discussion might be less accessible.
+Along the way, we examine the gauge redundancy of both the conjugate momentum (Section~\ref{gauge_momentum}) and the Lagrangian (Section~\ref{gauge}).
+
+%Despite being coordinate-invariant, one cannot make the density truly coordinate independent. We therefore argue that some quantities are physically significant despite being coordinate-dependent.
 
@@
 
-%[To do: Do we agree with the following claim defended by North: ``as far as we can tell, we need only symplectic structure to do classical mechanics" [74].
-If so, is our position in tension with the more general formulation of Lagrangian mechanics noted by Swanson and Halvorson, in terms of the almost tangent structure of the tangent bundle?
-[See their footnote 9, pages 4-5]
-
-In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by ``fundamental." Of the various notions of fundamentality, we will mean something entirely non-mysterious: a formalism is more fundamental provided that it is logically more general.
+In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by `fundamental.' Of the various notions of fundamentality, we will mean something entirely non-mysterious: a formalism is more fundamental provided that it is logically more general.
 In this case, the Lagrangian framework is less general because it requires that an additional condition hold, compared to the Hamiltonian framework.
@@
 Importantly, we require that the theory remains wholly \textit{about} classical mechanics.
-Hamiltonian mechanics is more general than Lagrangian mechanics while still being about classical systems, where ``aboutness'' can be understood using Yablo's framework \parencites*[]{Yablo}.
+Hamiltonian mechanics is more general than Lagrangian mechanics while still being about classical systems, where `aboutness' can be understood using Yablo's framework \parencites*[]{Yablo}.
 Having an account of \textit{aboutness} or \textit{subject matter} is a general philosophical problem that any account of fundamentality or essence must address.
 Hence, it is not a special problem facing our criterion.}
-% % What we really want is a brief formalization of ``proper subset," e.g. using partitions of possible worlds semantics. See if I can find Lewis discussion of this (maybe mentioned by Plebani)
-% % There is perhaps a further difficulty about the identity criteria for counting conditions.
-So we really just need it to be the case that all of the Lagrangian systems are a subject matter subset of the Hamiltonian systems.
-E.g.
-in the sense that the 1800s is a subset of all of history.
-need a mereological relationship.
-% Perhaps could say: this notion of a proper subset can be spelled out using Lewis on subject matter.
-Butterfield/gomes paper.
-But I could not find this discussion in their paper (might have a false memory)
-% % we could formalize this notion of ``proper subset" using what Lewis does regarding the mereology of partitions of possible worlds. This gives an account of subject matters, similar to Yablo's.
 
@@
 We thereby gain a straightforward interpretation of the physical significance of these formalisms.
-%Of the various notions of fundamentality, there is at least the following non-mysterious definition:
 
@@
 First, a system is \textit{infinitesimally reducible} provided that the state of the whole system is equivalent to specifying the states of its parts (and vice versa).
-Infinitesimal reducibility distinguishes classical from quantum systems, where the latter are irreducible (we cannot specify the whole state by specifying the states of its subsystems).
+Infinitesimal reducibility distinguishes classical from quantum systems, where the latter are irreducible (i.e.
+we cannot specify the whole state by specifying the states of its subsystems).
 In the widest sense, a classical system is any infinitesimally reducible system.
@@
 The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
-% % I think I could spend more time clarifying reducibility, infinitesimal. Maybe define separately first?
 
@@
 As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
-% % Would be ideal if we can provide some literature support for this notion of motion equivalence. E.g. connecting it with a discussion in Marsden and Abraham foundations of mechanics.
 
@@
 For at least this reason, Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
-% [To do: verify that the scenario of two different Lagrangians---giving rise to the same trajectories in physical space with different energies---arises ONLY when the motion equivalence condition fails.
-See Curiel 2014, page 42 for his take on this kind of scenario.]
-
-Of course, different definitions of ``Hamiltonian" and ``Lagrangian systems" can result in different interpretive relationships.
+
+Of course, different definitions of `Hamiltonian' and `Lagrangian systems' can result in different interpretive relationships.
 On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems satisfy the motion equivalence assumption.
@@
 Our derivation of Lagrangian mechanics in Section~\ref{motion} takes place within the context of conservative systems, so we do not consider systems that satisfy inhomogenous Euler--Lagrange equations.
-Although both the Lagrangian and Hamiltonian formulations apply to dissipative systems, we exclude these.\footnote{For details, see \parencites[\S 10.4]{Cline}.
+Although both the Lagrangian and Hamiltonian formulations apply to dissipative systems, we exclude these.\footnote{For details, see \textcites[\S 10.4]{Cline}.
 Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
@@
 In Wilson's terminology, the framework we apply here may be yet another `theory facade' for classical physics, but a fruitful one nonetheless.} This supports our contention that the property of infinitesimal reducibility fruitfully characterizes classical systems.
-Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by ``fundamental."
-%The property of infinitesimal reducibility provides a fruitful notion of a classical system.
-Classical systems are infinitesimally reducible systems.
-Hence, we feel this is the proper notion of a classical system.
-% For fluid systems, the deterministic/reversible evolution condition applies to the whole system but not for the individual particles due to interactions between these.}
-% % Why don't interactions between particles also arise for $n$-particle mechanics? e.g. Newtonian gravity?
-
-% % XXX: try to find a different citation for the claim that dissipative forces can be incorporated in both Hamiltonian and Lagrangian mechanics. Cline libre text might not look the best!
-
-% https://phys.libretexts.org/Bookshelves/Classical_Mechanics/Book%3A_Variational_Principles_in_Classical_Mechanics_(Cline)/10%3A_Nonconservative_Systems/10.04%3A_Rayleighs_Dissipation_Function
-
-%To avoid verbal disputes, it is essential to clarify what falls within the scope of ``classical systems." One way to delimit classical systems is to specify a set of equations that they satisfy---such as Hamilton's equations or the Euler--Lagrange equations.
-We will instead follow a different approach by delimiting classical systems based on the physical conditions that they satisfy.
-
-%Figure~\ref{diagram} summarizes our characterization of Hamiltonian and Lagrangian mechanics, in terms of physical conditions and corresponding mathematical structures.
-Section~\ref{derivation} justifies the logical connection between these physical conditions and corresponding mathematics.
-Before delving into this derivation, Section~\ref{fundamentality} discusses its philosophical significance.
-
-% One way to motivate the motion equivalence condition: in general, the same trajectory in physical space can correspond to different trajectories in phase space: knowledge of the position and velocity is not sufficient for knowledge of the momentum (e.g.
-if you do not know the mass of the system).
-Formally, different Lagrangians can have the same Lagrangian vector field but give rise to Hamiltonians with different Hamiltonian vector fields, as Barrett notes \parencites*[1186]{Barrett2}.
-In contrast, if two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to an arbitrary constant (whereas two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy, even up to a constant)).
-
-% %Overall summary: A classical system is an infinitesimally reducible system.
-The state space of these systems can be modeled by a symplectic manifold (the physical assumption of infinitesimal reducibility corresponds to the mathematical structure of symplectic manifolds).
-Classical Hamiltonian systems satisfy the additional conditions of having deterministic and reversible evolutions.
-These two physical assumptions entail the existence of a symplectic morphism on the state space manifold, making it a contact manifold.
-Classical Lagrangian systems satisfy a further assumption: keeping track of the trajectories in physical space is equivalent to tracking the evolution of states in phase space.
+Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by `fundamental.'
 
@@
 
-% North argues that ``the structure of physics is the (minimal) structure that is assumed among dynamically equivalent---in classical mechanics: canonically transformed---formulations of a theory.
-Structure comprises those quantities that remain intact when we transform one allowable coordinatization of the laws into another" [78]
-
-% North's epistemic method for figuring out the structure: ``look at what is invariant under allowable transformations"; this tells us what the structure is [73].
-
-As indicated at the start of Section~\ref{privileging}, we use ``fundamentality" in a non-metaphysical fashion.
-Differences in fundamentality are primarily differences in generality: more fundamental frameworks require fewer conditions to hold.\footnote{We also embrace related non-metaphysical uses of ``fundamental," discussed in Section~\ref{density}.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure." Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
+As indicated at the start of Section~\ref{privileging}, we use `fundamentality' in a non-metaphysical fashion.
+Some differences in fundamentality are differences in generality: more fundamental frameworks require fewer conditions to hold.\footnote{We also embrace related non-metaphysical uses of `fundamental,' discussed in Section~\ref{density}.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of `structure.' Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
 Compared with North's \parencites*[]{North} account of structure, our definition has two advantages.
@@
 Despite our different commitments, our approach has methodological parallels to North's.
-We agree with North that it is fruitful to identify both invariants and ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way" \parencites*[65]{North}.
+We agree with North that it is fruitful to identify both invariants and ``the mathematical structure needed to formulate the theory in an invariant, frame-independent way'' \parencites*[65]{North}.
 Additionally, we agree that it is worthwhile to identify what particular mathematical structure is required by a formulation \parencites[78]{North}.
@@
 See Section~\ref{Curiel} for discussion.} However, North advocates a metaphysical account of structure, about which we remain agnostic.
-According to North, ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description" \parencites*[66]{North}.
+According to North, ``structure comprises the objective, fundamental, intrinsic features, the ones that remain the same regardless of arbitrary or conventional choices and description'' \parencites*[66]{North}.
 These notions of fundamentality and intrinsicality correspond to differences in joint-carving, where more fundamental or intrinsic features are more joint-carving (or in Lewis's \parencites*[]{Lewis1983} terminology, more natural).
@@
 Precisely this scenario occurs in the context of Lagrangian mechanics, which requires the additional condition of motion equivalence to hold.
-Whereas North views Lagrangian mechanics as containing ``excess, superfluous structure" \parencites*[75]{North}, we take this additional condition as optional but still physically significant.
+Whereas North views Lagrangian mechanics as containing ``excess, superfluous structure'' \parencites*[75]{North}, we take this additional condition as optional but still physically significant.
 Thus, while it is illuminating to identify the minimum mathematical structures necessary for a formulation, additional structure is not automatically rendered physically insignificant.
-%It corresponds to the physical assumption of motion equivalence, which applies to many classical systems.
-%This characterizes the relationship between Hamiltonian and Lagrangian formalizations of a classical system. 
-% classical mechanics also deploys physically significant representations that are not intrinsic to mechanical systems [to do: mention some examples, e.g.
-in a footnote.
-The Lagrangian?
-Cartesian coordinates?].
-
-% Emphasize contrast: whereas North is arguing that we should believe that the state space of Hamiltonian mechanics corresponds to that of a classical world (using Occam's razor; inference to the best explanation), we are not trying to figure out which state space structure to believe in more.
-% As Swanson and Halvorson put it, North's argument presupposes ``that statespace structure actually corresponds to some kind of real, physical structure.
-Many physicists and philosophers of science view statespace as a formal tool for efficiently encoding certain modal commitments of scientific theories, and aid for making calculations and predictions, but not part of the theory's ontology.
-North's position presupposes some form of statespace realism" [1].
-North assuming that there is a fact of the matter regarding the state space of the world.
-
-Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real" fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure as part of the theory's ontology, but as a formal device for encoding physical information or as a calculational tool.
-As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is `fundamental in the bottom-layer sense' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental (in a classical world) does not matter for the questions we seek to answer.
-Regardless of its metaphysical status, Hamiltonian mechanics is physically more fundamental, and we will argue for its epistemic and conceptual primacy as well.
+
+Additionally, whereas North \parencites*[76]{North} seeks to determine the `real' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure ontologically, but as a formal device for encoding physical information or a calculational tool.
+As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is `fundamental in the bottom-layer sense' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental in a classical world does not matter for the questions we seek to answer.
+Regardless of its metaphysical status, Hamiltonian mechanics is physically and conceptually more fundamental.
 Indeed, our argument does not privilege any one of the various mathematical formalizations of Hamiltonian mechanics.\footnote{These include formulating Hamiltonian mechanics on a $2n+1$-dimensional contact manifold, on an extended phase space, on a Poisson manifold, and the standard formulation on a symplectic manifold that we typically use below.} Each of these frameworks has different pragmatic costs and benefits.
 Hence, it is plausible that there may be no single, uniquely necessary mathematical structure for mechanics, but rather a family of related structures.
-In other words, there may be no canonical minimal structure for formalizing the relevant physical models and physical assumptions.\footnote{Tulczyjew's reformulation of Hamiltonian mechanics supplies a further reason for viewing symplectic geometry---and hence Hamiltonian mechanics---as conceptually fundamental.
-According to Teh and Tsementzis, this reformulation shows that  ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh}.}  Additionally, each of these formalisms references some ``unphysical" coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.
-% We return to the physical significance of certain coordinate choices in Section~\ref{fundamental}.
-%The mathematical structure that you ``need" might not be unique.
-Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others.
-Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics.
-In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.
+In other words, there may be no canonical minimal structure for formalizing the relevant physical models and physical assumptions.\footnote{Tulczyjew's reformulation of Hamiltonian mechanics further supports viewing symplectic geometry---and hence Hamiltonian mechanics---as conceptually fundamental.
+According to Teh and Tsementzis, this reformulation shows that  ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry'' \parencites*[46]{Teh}.}  Additionally, each of these formalisms references some `unphysical' coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.
 
@@
 According to this criterion, statespace A has more structure than statespace B if the symmetry group of A is a proper subgroup of B's symmetry group \parencites[87-88]{North}.\footnote{This criterion follows from the most charitable reconstruction of North's argument, given by \textcites[]{Swanson}.
-North sometimes speaks in terms of comparing the dimensions of symmetry groups, which they note leads to counterexamples.} A symmetry group that is a proper subgroup entails that strictly fewer transformations preserve the space's structure.
+North sometimes speaks in terms of comparing the dimensions of symmetry groups, which they note leads to counterexamples.} Being a proper subgroup entails that strictly fewer transformations preserve the space's structure.
 This indicates that statespace A has more mathematical structure to be preserved.
@@
 Section~\ref{density} provides a further argument for viewing Hamiltonian mechanics as more fundamental.
-We argue that (i) invariant phase space densities are a fundamental physical quantity and (ii) Hamiltonian mechanics provides the proper setting for representing these densities.
-
-% % It is also a problem facing the metaphysical accounts of fundamentality, since the notion of ``essence" is also hyperintensional: e.g.
-Socrates is essentially human vs.
-Socrates is essentially (human and 2+2 = 4) [example from Stanford page].
-So we believe that we are in no worse shape than our more metaphysical opponents.
-
-%% Not sure if it is worth raising the following worry. If so, need to fix/clarify the response, which is sketchy.
-%Nevertheless, based on the equivalence between the tangent bundle and cotangent bundle formulations of Lagrangian and Hamiltonian mechanics, one might wonder whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics \parencites[1182]{Barrett2}.
-For, if there were, then this formulation of Lagrangian mechanics might be more general than the cotangent bundle formulation of Hamiltonian mechanics.
-And thus, there would be an analogous sense in which Lagrangian mechanics is more fundamental than Hamiltonian mechanics, undercutting our argument.
-Fortunately, it is implausible that there could be a suitable generalization of Lagrangian mechanics.
-In going from the cotangent bundle formulation of Hamiltonian mechanics to the symplectic manifold formulation, we lose some globally-defined structure.
-This is analogous to moving from the space of globally-Euclidean manifolds to the space of locally-Euclidean manifolds.
-Yet, it is not clear how Lagrangian mechanics could remain well-defined while throwing away global structure from the tangent bundle.
-For the action seems to be inherently a globally-defined quantity, as an integral defined over a trajectory.
-
-% Plausibly, move the following directly after our claim that our notion of fundamentality/structure comparison avoids the objections of Swanson/Halvorson against North's:
-
-% Way to motivate that the motion equivalence assumption further constrains the Hamiltonian: the Hamiltonian already defines a relationship between conjugate momentum and velocity (through ($dq/dt = \partial H / \partial p$) ).
-Motion equivalence provides a further constraint on this relationship, and hence a further constraint on the Hamiltonian.
-
-% As defined, our notions of fundamentality and structure are not metaphysical in nature.
-
-%Finally, our argument does not privilege any one of the at least three different mathematical ways of formalizing Hamiltonian mechanics [mention these three formalisms and provide references].
-Each of these frameworks has different pragmatic costs and benefits.
-Hence, unlike North, we are comfortable acknowledging that the mathematical structure that you ``need" might not be unique.
-There might not be a single canonical minimal structure for formalizing the relevant physical models and physical assumptions.
-Neither one of these three formalizations of Hamiltonian mechanics is necessarily more ``fundamental" or joint-carving than the others.
-And in some sense, each of these formalisms references some ``unphysical" coordinates or representational means.
-Nevertheless, there are compelling physical and pragmatic motivations for preferring the extended phase space formulation of Hamiltonian mechanics.
-In this formulation, we gain an invariant Hamiltonian that is the conjugate variable of the proper time, while the normal Hamiltonian (which is not coordinate-invariant) still plays the role of generating time-translations.
-% One con of this extended phase space formulation: the Hamiltonian constraint is both a constraint and the generator of time evolution; hence, it must be defined in the neighborhood of the actual states, so the values it provides outside of these neighborhoods should be unphysical, yet they encode the equations of motion.
-% Should we remark on the claim that the only structure we need is whatever is sufficient to define densities locally? Hence, we need just the symplectic form but not the cotangent bundle?
-% E.g.
-in the extended phase space formulation, the equations of motion come from derivatives of the Hamiltonian constraint over the variables, but this constraint is defined on unphysical points.
-These unphysical points are plausibly a mathematical artifact, although they supply a nice tool/method for understanding physically significant aspects of the representation.
-
-% One way to motivate the motion equivalence condition: in general, the same trajectory in physical space can correspond to different trajectories in phase space: knowledge of the position and velocity is not sufficient for knowledge of the momentum (e.g.
-if you do not know the mass of the system).
-Formally, different Lagrangians can have the same Lagrangian vector field but give rise to Hamiltonians with different Hamiltonian vector fields, as Barrett notes \parencites*[1186]{Barrett2}.
-In contrast, if two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to an arbitrary constant (whereas two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy, even up to a constant)).
-
-% Claim to think about: GC was arguing on 10-12 that you could handle classical systems in the Hamiltonian framework with just a differential manifold and no symplectic structure.
-This would contradict North's indispensability claim that Hamiltonian symplectic structure is necessary for treating classical systems.
-E.g., North footnote 31: ``The symplectic form is crucial to determining the motion, since the same Hamiltonian can induce different flows for different symplectic forms" [76].
-North takes this as an argument for why we cannot eliminate the symplectic form in an attempt to get at even less structure.
-We need it to define the dynamics.
+We argue that (i) invariant phase space densities are physically fundamental and (ii) Hamiltonian mechanics provides the proper setting for representing these densities.
 
@@
 \label{derivation}
-
-% [To do: figure out how to flag to the reader that they can skip/skim some parts of this derivation section without loss of understanding for later sections].
-E.g., `Readers who are more interested in the significance of invariant phase space densities or the geometric action principle can feel at peace skipping ahead.' But, presumably, part of the stuff on invariant densities is relevant for Section~\ref{density}?
-Figure out how to finesse this
 
@@
 
-% Is it worth emphasizing/highlighting any particularly important parts of the derivation up front?
-
 \subsection{Infinitesimal reducibility}
@@
 \begin{quotation}
-\noindent
+\bs \noindent
 \textit{Infinitesimal reducibility}: the state of the system is reducible to the state of its infinitesimal parts.
 That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the states of those parts, and so on.
+\es
 \end{quotation}
@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.
-For further discussion, see \textcites{shpmp}.
-[omit this for review].}
-%comment on how this characterization of infinitesimal particles relates to standard philosophy of physics interpretive disputes over the status of point particles in classical mechanics.
-Butterfield characterizes particles as extensionless.
-%For discussion of the status of extensionless point particles within classical mechanics, see [have yet to find decent reference for this kind of discussion].
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.}
+% % % For further discussion, see \textcites{shpmp}. [omit this for review].
 
@@
 If $\mathcal{S}$ is a discrete space with $U \subseteq \mathcal{S}$, we define $\mu(U) = \sum\limits_{s \in U} \rho_c(s)$ so that $\mu(\mathcal{S})$ gives us the size of the system and $\mu(U)/\mu(\mathcal{S})$ gives us the fraction of the particles that can be found in the region $U$.\footnote{The discrete case is far less interesting because discrete quantities cannot change continuously over time, meaning that they must be constant.
-For details see \textcites[]{AoPPhy1}.
-[omit this for review]}
-%[is region U in the state space C or S. I assume in S.].
+We discuss this elsewhere.}
+% % % For details see \textcites[]{AoPPhy1}. [omit this for review]}
 % % XXX could try to find justification/derivation for the state space being a manifold in the Kevin Kelly book, or could cite the math paper with Greenfield; proposition 6.2.
@@
 This density plays a crucial role throughout our argument.
-% A state variable change $\hat{\xi}$ represents the same state $s$ by a different set of state variables $\hat{\xi}^b$.
 
@@
 For convenience, we denote $\rho_{c, \hat{\xi}} (\phi(\xi^a))$ by $\rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
-Likewise, in the numerator of the Jacobian, we denote $\partial \phi$ by $\partial \hat{\xi}^b$, where here $\hat{\xi}^b$ is understood as a function of $\xi^a$.} Here, $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ is the determinant of the Jacobian for this change of state variables.
+Likewise, in the numerator of the Jacobian, we denote $\partial \phi$ by $\partial \hat{\xi}^b$, where $\hat{\xi}^b$ is here understood as a function of $\xi^a$.} Here, $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ is the determinant of the Jacobian for this change of state variables.
 From the preceding two equations, we see that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.
@@
 We will now show that in order for the density $\rho_{c, \xi}$ to have these transformation properties, the classical phase space must be at least a symplectic manifold.
-%footnote{Indeed, using the Lebesgue measure $\lambda$ on $\mathbb{R}^n$---and assuming that $\xi^{-1}$: $\mathbb{R}^n \to \mathcal{S}$ is measurable---we can define the measure $d\mathcal{S}$ by the pushforward measure $\xi^{-1}_{\ast}(\lambda)(U) \equiv \lambda(\xi(U))$.}
-%In other words, while we can change the units used by the state variables $\xi^a$ to identify the particle states, the units for $\rho^{\ast}_c$ must remain the same so that the density does not change its numeric value
 
@@
 Since the unit variables define the units for all state variables, this change must induce a unique change of state variables $\hat{\xi}^b = \hat{\xi}^b(\xi^a)$ for all variables.
-% [XXX: coordinates -> unit variables]
-
-% % Ask GC if instances of `state variables' in the next paragraph are correct; i.e. `state variables' rather than `coordinates'
-% On physical grounds, [XXX: I'd say it's more mathematical grounds]
-
-%This is because the distribution is a function only of the states, rather than of the particular choice of coordinates. 
- % Classical phase space (i.e. a symplectic manifold) is the only space that satisfies these constraints on the distribution. 
-%(in manifold terminology, these are known as coordinates)
 
@@
 We will show that, in fact, there must be exactly one $k^\alpha$ variable, conjugate to the single $q$ variable.
-% We further assume that $q$ fully defines the units for all state variables.
-[Is there a way to physically motivate this assumption, beyond this being a constitutive role of the position-like state variables?]
-
-As a scalar, the units of the density $\rho_c$ are invariant under changing the units of the state variables.
+
+As a scalar, the units of the density $\rho_{c, \xi}$ are invariant under changing the units of the state variables.
 So suppose we change the units by transforming $q$ to a new unit variable $\hat{q}=\hat{q}(q)$.
@@
 Therefore we must have two or more state variables, requiring at least one $k^\alpha$ variable.
-%[XXJ: if previously we measured the density in particles per meter cube---and we change units from meters to miles---won't the units of the density change to particles per cubic mile?
-Isn't this required by dimensional analysis on the expression $\int_U \rho^{\ast}_c(\xi^a) d\xi^a = \int_U \rho^{\ast}_c(\xi^b) d\xi^b$ ?].
-
-\begin{comment}
-%Longer justification for the Jacobian determinant equaling one.
-
- As defined, $\rho_c$ is a function from the particle state space $\mathcal{S}$ to the reals.
-We can define $\rho_c$ on the units by taking advantage of an invertible function $s: \mathbb{R}^n \to \mathcal{S}$ mapping a particular value of the units $\hat{\xi}^b$ to a point in the particle phase space $\mathcal{S}$.
-Hence, by definition $\rho_c(\hat{\xi}^b):= \rho_c(s(\hat{\xi}^b))$.
-[Is this function $s$ the inverse of the coordinate mapping $\xi^b$?].
-Then, since the physical state is independent of the coordinates, the state $s(\hat{\xi}^b)$ is the same state as $s(\xi^a) $ defined in terms of the original units $\xi^a$.
-Again, by definition, $\rho_c(s(\xi^a))=\rho_c(\xi^a) $.
-Since the distribution transforms as a density, it follows that $\rho_c(\xi^a)$ must equal the product of the Jacobian determinant $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| $ times the density defined over the transformed units: $\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)$.
-We can summarize the argument so far as a chain of equalities:
-
-\begin{equation*}
-\rho_c(\hat{\xi}^b)=\rho_c(s(\hat{\xi}^b))=\rho_c(s(\xi^a))=\rho_c(\xi^a) = \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| \rho_c(\hat\xi^b)
-\end{equation*}
-
-\noindent
-Therefore the Jacobian $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ must equal one.
-
-\end{comment}
 
@@
 Hence, the Jacobian is a lower triangular block matrix:
-
 \begin{equation}
@@
 So in the case of a single $q$ variable, defining an invariant density $\rho_{c, \xi}$ requires a two-dimensional manifold.
-
-\begin{comment}
-\begin{equation}
-\left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right| - \left|0\right| \left|\frac{\partial \hat{k}^b}{\partial q}\right|
-= \left|\frac{\partial \hat{q}}{\partial q}\right| \left|\frac{\partial \hat{k}^b}{\partial k^a}\right|.
-\end{equation}
- so $\left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right| = \left|\frac{\partial \hat q}{\partial q} \right|\left|\frac{\partial \hat{k}^b}{\partial k^a} \right|$: 
-\end{comment}
 
@@
 We then use $\hbar$ to define the units for state counting and therefore set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
-% % the `motivation' section of the wiki page on symplectic manifolds provides a different style of derivation for the 2-form being non-degenerate and closed.
 
@@
 We have shown that the state space of the particles must be a symplectic manifold.
-Additionally, any symplectic manifold structure supports the definition of coordinate-invariant densitys: any scalar function can be integrated using the symplectic form in a coordinate-invariant way.
-Therefore the symplectic structure is exactly the structure needed to define an invariant density.
+Additionally, any symplectic manifold structure supports the definition of coordinate-invariant densities: any scalar function can be integrated using the symplectic form in a coordinate-invariant way.
+Therefore symplectic structure is exactly the structure needed to define an invariant density.
 Section~\ref{density} describes further interpretive consequences of this connection between invariant densities and symplectic structure.
@@
 %Symplectic manifold structure is exactly the structure needed to define an invariant density.
-% Could connect this with the earlier discussion of North's methodology of seeking coordinate-invariants, from section~\ref{fundamentality}.
-
-%since the degrees of freedom are independent, the number of states on a surface does not change if we translate that state across independent degrees of freedom.
-If we imagine a parallelepiped, the integral over the surface must be zero (the integrals over opposite sides are equal and opposite) [explain/justify this more clearly.
-Is this coming from the two-form being antisymmetric?].
-%
-%if degenerate form, would have states that are not counted by the form.
-
-% Which is the only type of manifold that supports coordinate-invariant distributions
 
@@
 \begin{quotation}
+\bs
 \noindent
-\textit{Deterministic and reversible evolution}: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified.
-
+\textit{Deterministic and reversible evolution}: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified. \es
 \end{quotation}
@@
 the density remains the same throughout the evolution.\footnote{For convenience, we will sometimes drop the composite state $c$ and coordinate-choice $\xi$ indices from the invariant density $\rho_{c, \xi} (\xi^a)$.
-The fact that the density is not coordinate-free is central to our argument in Section~\ref{density}.
 We will reserve the word `distribution' for the coordinate-free distribution $\rho_c (s)$.
 } In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa.
-%[could two different particles occupy the same initial state?].
-%[what exactly does `initial particle' mean? Do we mean the initial state of the composite system or just a particular particle?]
+%The fact that the density is not coordinate-free is central to our argument in Section~\ref{density}.
 
@@
 Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes (i.e.
-sets of states are mapped to sets of states of equal size) which leads to Hamiltonian evolution.
+sets of states are mapped to sets of states of equal size), which leads to Hamiltonian evolution.
 In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.
@@
 
-% [to do: make the footnote regarding connection to Marsden and Abraham more precise.
-Might want to flag that their theorem is expressed in terms of the Lagrangian vector field and corresponding Hamiltonian vector field (obtained via Legendre transformation) having the same base integral curves.
-These base integral curves encode how particles evolve over time.
-Should I also elaborate on how one can interpret motion equivalence in terms of the theories having the same empirical content?
-This is how Barrett interprets the Marsden and Abraham theorem 3.6.2]
 \begin{quotation}
-\noindent
+\bs \noindent
 \textit{Motion equivalence assumption}: the motion of the system (i.e.
@@
 evolutions in state space) and vice-versa.\footnote{ \textcites[218]{Abraham}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time; see \textcites[1180-1181]{Barrett2} for discussion.
-Motion equivalence can be understood as the converse of this theorem: assuming that Lagrangian and Hamiltonian mechanics agree on particle evolutions, we derive that the Legendre transformation is well-defined in the hyper-regular domain.}
+Motion equivalence can be understood as the converse of this theorem: assuming that Lagrangian and Hamiltonian mechanics agree on particle evolutions, we derive that the Legendre transformation is well-defined in the hyper-regular domain.} \es
 \end{quotation}
@@
 
-%<<<<<<< Updated upstream: XXJ: wasn't sure what this referred to? 
 In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^i$.
@@
 Let $H$ denote the Hamiltonian.
-By Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
+By Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, so the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
 Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere---which we take to be positive.
@@
 Note that while the value given by $\rho(q^i, p_j)$ is invariant under unit changes, the value given by $\rho(x^i, v^j)$ depends on the choice of units through $\left|\frac{\partial v^i}{\partial p_j}\right|$.
-If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only, since the position variables---by themselves---define the units.
+If $x^i$ truly sets the unit system by itself, then $\left|\frac{\partial v^i}{\partial p_j}\right|$ must be a function of position only, since the position variables define the units.
 Similar considerations hold for marginal distributions (i.e.
 distributions on a subset of the unit variables), entailing that all components of $\frac{\partial v^i}{\partial p_j}$ must be a function of position only.
-We set:\footnote{In Section~\ref{dependence}, we will see that $g^{ij}$ corresponds to the Riemannian metric, since $\partial v^i/\partial p_j$ is a function only of position.}
+We set:
+%\footnote{In Section~\ref{dependence}, we will see that $g^{ij}$ corresponds to the Riemannian metric, since $\partial v^i/\partial p_j$ is a function only of position.}
 \begin{equation}
@@
 
-\begin{comment}
-* ``Fundamentality" because it makes the physics more prominent.
-Some people use coordinate-free/independence as necessary conditions for fundemantaly/physical significance.
-This means coordinates do not appear in the definition.
-Yet, densities require coordinate for the definition, therefore cannot be coordinate-free/independent, therefore there are some physically significant quantities that do not satisfy that.
-Why fundamental???
-(reverse mathematics?
-fundamental because it recovers the correct math from the physical assumption?)
-
----
-Section on densities and riemmanian geometry. (Gabriele)
-
-* A paragraph to show how densities work.
-In general are coordinate dependent and you can't simply compare values at different points.
-This is the case in Lagrangian mechanics because both differentials in position vary contravariantly.
-
-* Contrast this with densities over position and velocity.
-But!
-We can understand some aspects of Riemmanian geometry in terms of how states densities are represented over position and velocity.
-Hyper-regularity allows us to express densities from "states" to "kinematic variables".
-The fact that this expression depends only on position (and not on velocity), tells us we have a Riemannnian manifold.
-Norm of the metric tensor gives us the ``conversion'' factor.
-Note that norm of the metric tensor in general depends on the coordinate system and is in general point-dependent.
-There always exists a coordinate system in which the norm of the metric tensor is constant.
-The conversion factor for marginal distributions is linked to the eigenvalues along each coordinate.
-Flatness means there exists a coordinate system where we can factorize the distribution (if it factorizes in phase space), which is exactly the Cartesian coordinate system.
-(Note, we can always factorize in phase space) -> Therefore this focus on densities also allows us to get a better intuition even on the tangent space.
-[Still need to incorporate much of this discussion]
-
-\end{comment}
-
 Hamiltonian mechanics is more fundamental than Lagrangian mechanics because Hamiltonian systems require fewer physical conditions.
@@
 In general, it provides a better understanding of classical systems.\footnote{It also provides a better understanding of statistical mechanics and entropy, since phase space provides a uniform measure over q and p.
-For discussion, see \textcites{shpmp} [omit for review].}
+We discuss this elsewhere.}
+% % % For discussion, see \textcites{shpmp} [omit for review].}
 
@@
 Typically, the canonical position and momentum observables play this role, since they suffice to completely specify a classical state; they fix the values of all other observables.
-\textcites[31, 200]{Ruetsche} refers to this as ``fundamental in the physicist's sense." However, we will show in Section~\ref{gauge_momentum} that the conjugate momentum is gauge-redundant.
+Ruetsche refers to this as ``fundamental in the physicist's sense'' \parencite*[31, 200]{Ruetsche}.
+However, we will show in Section~\ref{gauge_momentum} that the conjugate momentum is gauge-redundant.
 For this reason, the invariant density provides a more fundamental specification of the state than the state variables themselves.\footnote{Section~\ref{infinitesimal} provides another reason for viewing the density as fundamental: by defining the density $\rho_{c, \xi}$, we recover the right mathematical structure, namely a symplectic manifold.
@@
 The syntactic symmetry between position and momentum in the mathematical formalism does not entail that these quantities are equally physically significant.
-%\footnote{This illustrates how one can be mislead by the syntactical form of the mathematics to draw an interpretive conclusion that is not borne out by the physics.}
-
-% % perhaps put at end of section 5.1: We have just shown that Hamiltonian mechanics provides the proper setting for representing the density.
-By representing the density on a symplectic manifold, we do not need to know which coordinate system we have chosen to work in.
-
-% Canonical coordinates are physically significant because they allow us to define an invariant phase space density.
-% % This latter claim might seem puzzling, since insofar as we need to use canonical coordinates to define the density, this seems to speak in favor of the canonical coordinates being more fundamental.
-
-\subsection{Coordinate-dependence and coordinate-variance}
+
+\subsection{Point-wise comparison}
 \label{dependence}
-% densities as coordinate -dependent and variant
-
-%It is helpful to take a quick look at how densities work in general so that we can better understand what is so special about them in the context in Hamiltonian mechanics.
 
@@
 Or we may represent these characteristics using height $h$ and body mass index $bmi = w / h^2$, which leads to a different density $\rho(h,bmi)$ with units of (people$\cdot$m)/Kg.
-
-% In polar coordinates $(r, \theta)$, if we measure $r$ in meters and $\theta$ in radians, then $\rho(r, \theta)$ must return mass over meter-radians.
-[changed example.
-Use height/weight to height/body mass index.
-Issue with geometrical examples is that one automatically thinks of geometrical volume.]
 
@@
 
-%there is more stuff at one location in state space than another, we generically need to know which coordinate system we are in, and how the density depends on these coordinates (i.e.
-we can't just compare the values of $\rho$ in these coordinates and see which is larger).
 %The Hamiltonian formalism circumvents this further epistemic requirement by defining the density using quantities that vary covariantly.
@@
 when we express it over a manifold, then it needs to be a density for integration, and hence it becomes coordinate-dependent).
-[see 57.3 minutes of 12-21 recording for discussion]
 
@@
 
-%The gauge-dependency of the conjugate momentum supplies further grounds for privileging the invariant phase space density (as the fundamental description of a classical state)?
-
 Whereas we privilege the invariant phase space density, some philosophers have instead privileged the canonical state variables $(q, p)$.
-For instance, in her argument for the fundamentality of Hamiltonian mechanics, North claims that ``momentum is a truly fundamental quantity, on a par with position" \parencites*[77]{North}.
+For instance, in her argument for the fundamentality of Hamiltonian mechanics, North claims that ``momentum is a truly fundamental quantity, on a par with position'' \parencites*[77]{North}.
 Likewise, \textcites[164,171]{Wallace} claims that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
@@
 Instead, we should understand the conjugate momentum $p$ as a label that we can use to express the invariant density over states.
-In this way, the invariant density is more physical than the state variables themselves.
+In this way, the invariant density is more fundamental than the state variables themselves.
 % % after Double checking Wallace citation, I should probably find a better/more direct one. e.g. Earman article in that volume?
-% Possibly a stronger claim: This entails that the invariant density is really the important physical element/object in Hamiltonian mechanics.
-% % Could also connect standard interpretation of the canonical state variable pair with the earlier quotation from Ruetsche, as these being fundamental in the physicist sense.
-Shows how we are providing a more fine-grained physical notion of fundamentality, by going beyond just the epistemic criterion that Ruetsche provides.
-Both the invariant density and the canonical state variable pair are fundamental in the physicist sense.
-But the invariant density is more fundamental because it is not gauge-dependent.
 
@@
 We will show that these transformations are exactly the gauge transformations.
-
-% To show that the conjugate momentum is gauge-dependent, consider the set of transformations that leave the density, position, and velocity invariant.
-We will see that we can interpret these transformations as a set of gauge transformations on the conjugate momentum $p $.
-This entails that the conjugate momentum is a gauge-dependent quantity, unlike both the position state variable $q$ and the \textit{kinetic} momentum.
-Recall that motion equivalence entails the following: knowledge of the trajectory in physical space suffices for knowledge of the state in phase space.
-In other words, we can completely characterize states in terms of the canonical position coordinate $q$ and its first derivative $\dot{q}$.
-Hence, any transformation that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant.
-We will further assume that the transformations leave the density invariant, which entails that the new variables must be canonical.
 
@@
 Insofar as coordinate-invariance is a plausible necessary condition for fundamental physical status, so is gauge-invariance.
-By being gauge-dependent, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can remain physically significant but not fundamental).
-%[XXX: still, conjugate momentum is less physically significant because is not well defined experimentally]
+By being gauge-dependent, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can be physically significant but not fundamental).
+%might not need this paragraph; tacitly references the section 5.2 I deleted.
 
@@
 It is in this sense that the unit-variables $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.
-Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are of equal physical significance.
+Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are equally significant.
 In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
@@
 %\footnote{As discussed in Section~\ref{fundamental}, these units are physically significant despite not being coordinate-invariant.}
-
-% Whereas symplectic geometry creates a mathematical, syntactic symmetry between the $q$ and $p$ variables, this is not a physically significant symmetry from the standpoint of measurement.
-%cite van Fraassen on the measurement point; e.g. APA session on Halvorson book
-
-% % [To do: explain why we have grounds to privilege the invariant density over not only the conjugate momenta but also the canonical position coordinate.
-Is this connected with the following thought: Hamiltonian mechanics does not really care what the q's and p's are.
-It is the motion assumption that tells us that q's are position, simply because they define the unit system.
-]
-% Possible response: the invariant density by itself can specify the state of the classical system.
-Whereas the canonical position coordinate (although fundamental in its own right) is insufficient for defining the state.
-So it is only the pair of canonical state variables that could define the state, but we have problematized the fundamental status of the conjugate momentum.
-
-% % [To do: think about whether treating gauge-invariance as a necessary condition for fundamentality is in tension with our position, specifically our contention that coordinate-dependency does NOT preclude a quantity from fundamental physical significance.
-Does our position require an important physical difference between gauge-dependency and coordinate-dependence?].
-Maybe the thought is that gauge-invariance and coordinate-invariance are on the same footing.
-So if coordinate-invariance is a necessary condition for physical significance, then so is gauge-invariance.
-A gauge-dependent quantity can still be physically significant, in the same way that a coordinate -dependent quantity can be physically significant.
-% % Note that if coherent, our position has the following consequence: gauge-dependency is a sufficient condition for a quantity to be non-fundamental (in the same way that coordinate-variance is sufficient for non-fundamentality).
-But coordinate-dependency is NOT a sufficient condition for being non-fundamental.
-This could strike some readers as in tension.
-If we are comfortable viewing gauge-dependency as precluding a quantity from fundamental significance, then why not coordinate-dependence?
-Response: we just need to motivate that gauge-dependency is more analogous to coordinate-variance than coordinate-dependence.
-
-% Have mostly done the following: [To do: expand on the following point, and try to eliminate the notion of ``simplicity" and replace it with a more epistemic notion.
-Ability to make comparisons of density, independently of which coordinate system you are in or knowledge of being in a ``good" coordinate system.
-Josh: see 11-30-20 meeting notes where I think we discussed this.]: Compared with Lagrangian mechanics, canonical pairs (position and conjugate momentum) allow a simpler comparison of densities over space.
-
-% [Not sure if/where the following claim is relevant: ] This difference in the number of assumptions corresponds to a difference in the number of commitments.
-When working with a Hamiltonian but non-Lagrangian structure, we are not committed to the q's being positions (although is this in tension with the remarks on the significance of the coordinate and unit systems?).
-In contrast, for systems where motion equivalence applies, we gain the further commitment of treating the q's as positions.
-It is the motion equivalence assumption that tells us that q's are position, simply because they define the unit system.
-P is a gauge quantity (we care about the trajectory, the velocity).
-
-% By definition, a dynamical system is one where we can write xdot as a function of x (i.e.
-the current state determines the evolution of the system).
-Having a Hamiltonian system further means that we can express this function xdot(x) using the Hamiltonian.
-
-% We took out the following because coordinate-invariance in fact did not seem to be either a necessary or sufficient condition for physical significance: Assuming that coordinate-invariance is sufficient for physical significance, the density thereby represents a genuine physical property of classical systems.
 
@@
 We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a related argument, see \textcites[236--37, 243--45]{Arnold}.}
-%our derivation seems very similar to Arnold chapter 9, section 45:C: ``the principle of least action in phase space''. Note that we briefly discussed Arnold on 11-23-20 meeting, around 12:30. 
-% Note on the generalization: cannot use curl.
-Need to use a form that measures the number of states on a surface.
-The displacement field kills this form (eats the degenerate direction that does not form degrees of freedom).
-% Something that I thought was a second assumption but turns out that this follows just from double role played by time variable: 2) the system's motion always advances in time (so that $dt/dt = 1$).
 
@@
 We could instead use a parameter $\tau$, and the derivation would remain valid.} This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $.
-In components, this means that $S_t = - (\frac{\partial}{\partial q} \cdot \theta_p - \frac{\partial}{\partial p} \cdot \theta_q) = - (\frac{\partial}{\partial q} \cdot 0 - \frac{\partial}{\partial p} \cdot \theta_q) = \frac{\partial q}{\partial p}$.
-Hence, $S_t = 1$ entails that $\frac{\partial q}{\partial p} = 1$.
+In components, this means that $S_t = - (\frac{\partial}{\partial q} \cdot \theta_p - \frac{\partial}{\partial p} \cdot \theta_q) = - (\frac{\partial}{\partial q} \cdot 0 - \frac{\partial}{\partial p} \cdot \theta_q) = \frac{\partial \theta_q}{\partial p}$.
+Hence, $S_t = 1$ entails that $\frac{\partial \theta_q}{\partial p} = 1$.
 Integrating this expression, we find that $\theta_q = p$ up to addition by an arbitrary constant function $g(q)$.
@@
 These simply are Hamilton's equations, which justifies interpreting $H$ as the Hamiltonian.
-
-% Since the system is deterministic and reversible, it satisfies Hamilton's equations: $\frac{\partial H}{\partial p} =\frac{dq}{dt} \equiv S_q$ and $-\frac{\partial H}{\partial q} =\frac{dp}{dt} \equiv S_p $.
-Additionally, since $\vector{S} = -curl(\vector{\theta}) $, we know that $S_q = -\frac{\partial \theta_t}{\partial p}$ and $S_p = \frac{\partial \theta_t}{\partial q}$.
-Integrating the first equation for $\theta_t$ gives $-H$ up to an arbitrary constant function of q, while the second gives $\theta_t = -H$ up to an arbitrary constant function of p.
-For consistency, these two constant functions must vanish, so that $\theta_t = -H$ [check argument here; do we actually need to appeal to the gauge freedom of S again?].
-Hence, it follows that we can take $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
 
@@
 To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
-  %Provided that the Legendre transform is well-defined (i.e.
-the motion equivalence assumption is satisfied), then $H = p \dot{q} - L$.
-Hence, we can re-express the action as $\integral^{t_1}_{t_0} L (q, \dot{q}, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$.
-% Note on the generality of the argument: in cases where motion equivalence is not satisfied, we can still write the integrand in terms of H(q, p, t), although it would not be an integral over the Lagrangian.
-Thus, we can still express the `action' as a function of p, q, t.
-We would calculate this on a path in the extended phase space, although without motion equivalence, not all of the paths on phase space would correspond to a unique trajectory in real, physical space.
-See if Arnold has a more general name for this generalized principle of stationary action.
 
@@
 Hence, by Stoke's theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot dC = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ .
-Recalling now that i) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and ii)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $curl(\vector{\theta})$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
+Recalling that i) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and ii)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $curl(\vector{\theta})$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
 This establishes a more general Hamilton's principle of stationary action, which we have derived simply by assuming that the displacement vector is divergenceless.
@@
 
-%(XXX to do (might no longer be necessary given earlier and later remarks) Note how the geometry is better understood in extended phase space, requiring the Hamiltonian picture.
-Note that this applies even when $\theta \cdot S$ cannot be expressed as a function of $q$ and $\dot q$.
-)
-
-%Discussion of gauge transformations in the context of Lagrangian vs.
-Hamiltonian mechanics; contrast with North [77] privileging the momentum variables as significant (look for Wallace 2003 quotations along these lines as well):
-
-\subsection{Gauge redundancy in Lagrangian mechanics}
+\subsection{Gauge redundancy of the Lagrangian}
 \label{gauge}
-
-\begin{comment}
-
-% 12-19: I think that I incorporated the relevant points below:
-
-** Overall point: this approach allows us to get a better intuition for what gauge transformations represent.
-
-* From motion equivalence -> we really only care about the trajectory. That is what we want to represent and what we measure.
-
-* A gauge transformation is a change in p variable that leave q and q dot unchanged.
-Therefore it is simply a change of how states are labeled, without a change in the actual observable dynamics.
-Add the derivation.
-
-* In Hamiltonian mechanics, the momentum is gauge dependent.
-In Lagrangian mechanics the kinematic variables are not gauge dependent, but the Lagrangian manifestly is.
-$L = \theta S$.
-$S$ is gauge independent but $\theta$ is not.
-Therefore we will have problems understanding the physicality of the Lagrangian.
-Similar concerns for the Action.
--> neither formulation has the upper hand
-
-* Coordinate transformation: if we restrict to coordinate transformation that only depend on space, both Lagrangian and Hamiltonian are invariant.
-If we introduce coordinate transformation that do depend on time (e.g.
-rotating frame), then Lagrangian is still invariant while Hamiltonian manifestly changes as a time component of a co-vector.
-(because H is the time component of $\theta$).
-That is why the Lagrangian framework is prevalently used in intro classes.
-
-* This is a problem of the double role of time as ``coordinate'' and ``affine parameter''.
-To be proper in differential geometry, we have to separate the two uses.
-So we add a ``temporal'' degree of freedom $(t, E)$ and the evolution is over the affine parameter $s$.
-In that framework, the Hamiltonian constraint serves both as a global constraint that fixes the energy as a function of $q, p, t$ and as the generator of the affine parameter $s$.
-The Hamiltonian constraint is invariant.
-If the affine parameter is proper time, then the Hamiltonian constraint is essentially the mass.
--> There is a better formulation of Hamiltonian mechanics, that is typically not found in textbooks, that addresses the problem.
-
-\end{comment}
-
-% Last page of Halvorson/Swanson article seems to broach this issue of gauge redundancy, although they seem to think it is possibly a positive feature of Lagrangian mechanics (a source of additional representational flexibility) rather than a defect.
-
-% Understanding mechanics in terms of our assumptions gives us a way to understand gauge transformations that is both more physically grounded and intuitive.
 
@@
 In the next section, we rebut Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.
-%%Arnold uses a related extended phase space $(q_i, p_i, t)$ in his chapter 9, but doesn't provide a general discussion of it. 
-%Discussion by GC of extended phase space: https://tinyurl.com/4tlbej76
-% Is the Hamiltonian constraint for the extended phase space the same as the discussion in this Wikipedia page: https://en.wikipedia.org/wiki/Hamiltonian_constraint
-
-%[To do: not sure how best to end this subsection/transition to the next one].
-
-\begin{comment}
-
-I think that I have incorporated most of the following into the above:
-
-I didn't incorporate the following point: In some sense (to be made more precise), gauge transformations are more prevalent in Lagrangian mechanics than Hamiltonian mechanics.
-This is so despite the fact that the conjugate momentum $p $ is itself a gauge quantity.
-
-% [To do: clarify the following claim:] motion equivalence allows us to understand the gauge redundancy of the Lagrangian.
-However, it is not true that the Lagrangian has these redundancies in virtue of the system satisfying motion equivalence.
-
-[To do: clarify the following :] From motion equivalence, the Lagrangian has more invariant structure than the Hamiltonian [not sure what this really means].
-The Lagrangian has the advantage of being invariant under coordinate transformations, whereas the Hamiltonian is not a coordinate invariant.
-Instead, the Hamiltonian varies as the time component of a 4-co-vector.
-Nevertheless, the Lagrangian itself is physically problematic because it is a gauge-dependent quantity [elaborate/illustrate this].
-
-[Clarify the following, if relevant :] But there is a sense in which on the extended phase space, the Hamiltonian constraint is an invariant, yielding a coordinate-invariant way of handling systems with Hamiltonian mechanics.
-
-[To do: connect the following point with the former :] Symplectic geometry indeed does not care about the difference between q's and p's.
-Classical mechanics does.
-Misindentification of the math (symplectic geometry) with the physics (classical Hamiltonian mechanics).
-%Since the q and $\dot{q} $ variables fix the physical state, this indicates that transformed conjugate momentum is a gauge quantity. 
-%North argues that ``momentum is a truly fundamental quantity, on a par with position" [77]. I think that Wallace 2003 article might make a similar point. Look for this.
-
-\end{comment}
-
-\begin{comment}
-Issue of the mathematical structure that is necessary/sufficient for Lagrangian mechanics:
-North footnote 20 and appendix: it is not generically true that the state space of Lagrangian mechanics comes equipped with a metric.
-Footnote 22: ``a more general, coordinate-free version" of Lagrangian mechanics ``does not make explicit use of this metric"; discussed in appendix A.
-North: ``the square of the Riemannian line element is what gives the allowable sets of coordinate transformations, the ones that preserve the Lagrangian equations of motion.
-A volume element would not suffice for this" [75].
-How does this connect with motion equivalence and its associated mathematical structure?
-
-North assumes that formulating a Lagrangian theory requires specifying a Riemannian metric [but this is not true: see the discussion of her appendix A above]; according to Curiel, this is neither necessary nor sufficient.
-Furthermore, the physical significance of a Riemannian metric in the context of a Lagrangian theory is unclear: two different Lagrangians can have the same Lagrangian vector field and yet induce inequivalent Riemannian metrics
-
-%Somewhere in this section, should reply to Curiel's privileging of Lagrangian mechanics.
-Curiel frequently appeals to naturalness and a distinction between natural vs.
-ad hoc/artificial constructions [he defines `naturalness', specifically a `natural coordinate system', in terms of a particular kind of coordinate transformation, using the notion of the almost tangent $J $ structure.
-See his equation 4.3, page 22 of the PDF].
-
-% Helpful discussion of differences between Riemannian and symplectic manifolds from Wikipedia (https://en.wikipedia.org/wiki/Symplectomorphism): ``Unlike Riemannian manifolds, symplectic manifolds are not very rigid: Darboux's theorem shows that all symplectic manifolds of the same dimension are locally isomorphic.
-In contrast, isometries in Riemannian geometry must preserve the Riemann curvature tensor, which is thus a local invariant of the Riemannian manifold.
-Moreover, every function H on a symplectic manifold defines a Hamiltonian vector field XH, which exponentiates to a one-parameter group of Hamiltonian diffeomorphisms.
-It follows that the group of symplectomorphisms is always very large, and in particular, infinite-dimensional.
-On the other hand, the group of isometries of a Riemannian manifold is always a (finite-dimensional) Lie group.
-Moreover, Riemannian manifolds with large symmetry groups are very special, and a generic Riemannian manifold has no nontrivial symmetries."
-%Seems to me that this theoretical result supports North's analysis, and could offer a response to Halvorson/Swanson counterexamples that appeal to Kepler problem and Galilean boosts.
-
-\end{comment}
 
@@
 
-% Helpful summary of Curiel's position: ``Hamiltonian mechanics does not respect the kinematical constraints intrinsic to abstract classical systems" while ``Lagrangian mechanics always respect[s] the kinematical constraints of abstract classical systems" \parencites*[304]{Curiel}.
+% Helpful summary of Curiel's position: ``Hamiltonian mechanics does not respect the kinematical constraints intrinsic to abstract classical systems'' while ``Lagrangian mechanics always respect[s] the kinematical constraints of abstract classical systems'' \parencites*[304]{Curiel}.
 
@@
 \end{equation}
-Curiel contends that this form is the ``necessary and sufficient condition [a Hamiltonian must have] for it to model an abstract classical system" \parencites*[305]{Curiel}.
-Since many possible Hamiltonians do not take this form, Curiel believes we must artificially restrict Hamiltonian mechanics to those that satisfy Equation~\ref{6.2}.\footnote{As an example, Curiel considers the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$  \parencites*[305]{Curiel}.}
-%In contrast, Curiel believes that Lagrangian mechanics does not require an analogous restriction.
+Curiel contends that this form is the ``necessary and sufficient condition [a Hamiltonian must have] for it to model an abstract classical system'' \parencites*[305]{Curiel}.
+Since many possible Hamiltonians do not take this form,\footnote{As an example, Curiel considers the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$  \parencites*[305]{Curiel}.} Curiel believes we must artificially restrict Hamiltonian mechanics to those that satisfy Equation~\ref{6.2}.
 
@@
 Moreover, our interpretation remains compatible with the claim that many---perhaps even most---non-regular Hamiltonian systems are physically uninteresting.
-%could footnote Curiel pages 309-310 discussion of Hamiltonians for systems with constant velocity. see longer comment note below.
-
-Whereas we can easily interpret non-hyper-regular Hamiltonians, the same cannot be said for all non-hyper-regular Lagrangians.
+
+Whereas we can easily interpret non-hyper-regular Hamiltonians, the same cannot be said for non-hyper-regular Lagrangians.
 For example, consider the case where the Lagrangian $L$ equals zero.
-For this ``system," all paths have the same action, so all paths are possible.
+For this `system,' all paths have the same action, so all paths are possible.
 Given an initial probability distribution over states, the Lagrangian fails to specify how this distribution will evolve in time.
@@
 a unique evolution in phase space) even when outside the hyper-regular domain (i.e.
-when the Hessian is zero).\footnote{In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
-There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.} For instance, a zero Hamiltonian describes a static system (by Hamilton's equations, the derivatives of the position and momenta are zero).
+when the Hessian is zero).\footnote{Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
+There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other'' \parencites*[1179]{Barrett2}.} For instance, a zero Hamiltonian describes a static system (by Hamilton's equations, the derivatives of the position and momenta are zero).
 Indeed, Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside the hyper-regular domain \parencites*[291, 308]{Curiel}.
-Curiel argues that this is an advantage of Lagrangian mechanics, but it generates challenges for understanding non-hyper-regular Lagrangians.\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.
+Curiel argues this favors Lagrangian mechanics, but it generates challenges for understanding non-hyper-regular Lagrangians.\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.
 In the case where a system is non-regular only at certain points, we can decompose the system into hyper-regular subsystems, each with its own Lagrangian (or Hamiltonian).}
-%The non-hyper-regular points form a boundary that we disregard.
-
-%but he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage.
-On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}.
-In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy the proper kinematical constraints (e.g.
-$v^i = \dot{x}^i$).
-Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will allegedly still satisfy these kinematical constraints \parencites*[311]{Curiel}.
-Curiel therefore views these non-regular, Lagrangian but non-Hamiltonian systems as physically-meaningful classical systems.
 
@@
 Preceding sections have illustrated many of these advantages.
-Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{density}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a more geometrical and general understanding of the principle of stationary action (Section~\ref{action}).
+Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{dependence}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a general geometrical understanding of the principle of stationary action (Section~\ref{action}).
 Additionally, our definition accommodates models of classical massless particles by maintaining the uniqueness of solutions, even outside the hyper-regular domain.
-Moving beyond classical mechanics, our framework has the advantage of providing an elegant relationship with quantum mechanical systems.\footnote{For details of the quantum case, see \textcites{AoPPhy1}.
-[omit for review].} By changing the requirement of infinitesimal reducibility (Section~\ref{infinitesimal}) to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.
-Hence, our framework perspicuously distinguishes quantum from classical systems.
+Moving beyond classical mechanics, our framework has the advantage of providing an elegant relationship with quantum mechanical systems.
+By changing the requirement of infinitesimal reducibility to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.\footnote{We provide details of the quantum case elsewhere.} Hence, our framework perspicuously distinguishes quantum from classical systems.
 We take these considerations as reason enough for adopting our preferred notion of a classical system.
-%Both kinds of systems are Hamiltonian thanks to deterministic and reversible evolution.
-
-%On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems).
-In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers the Hamiltonian $H=p$, which corresponds to a system moving with constant velocity (e.g.
-velocity of 1).
-So this could very well be a model of a photon, moving at the speed of light.
-Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating the Euler--Lagrange equation.
-Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310].
-But then does Curiel consider this system to not be a classical system?
-Additionally, insofar as this corresponds to the trivial Lagrangian, does this pose a problem for his own endorsement of Lagrangian mechanics outside the hyper-regular domain?
-% in the model of the classical photon, it clearly has a velocity.
-e.g., you always have velocity in Hamiltonian mech.
-one law tells you how velocity is linked to position and velocity.
-So it is not clear what curiel means by saying that non-regular hamiltonian system would not satisfy this kinematic constraint.
-
-% Another reason to privilege our definition: our state space works well for defining the density.
-Whereas using position/velocity leads to the problems with the density that we saw in the preceding section.
-Curiel's framework works well for systems that have inertial frames, Cartesian coordinates, and are closed rather than open.
-Position/velocity variables work well for systems like these, but not if you break one of these conditions.
-In our framework, we have the disadvantage of the gauge dependence of momentum, but this is a small price to pay.
-
-\begin{comment}
-
-% Material that I cut from the Curiel section:
-
-%[To do: briefly expand on this and connect to Curiel page 309-310 discussion of a similar system] 
-% in the model of the classical photon, it clearly has a velocity.
-e.g., you always have velocity in Hamiltonian mech.
-one law tells you how velocity is linked to position and velocity.
-So it is not clear what curiel means by saying that non-regular hamiltonian system would not satisfy this kinematic constraint.
-
-%On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems).
-In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers the Hamiltonian $H=p$, which corresponds to a system moving with constant velocity (e.g.
-velocity of 1).
-So this could very well be a model of a photon, moving at the speed of light.
-Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating the Euler--Lagrange equation.
-Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310].
-But then does Curiel consider this system to not be a classical system?
-Additionally, insofar as this corresponds to the trivial Lagrangian, does this pose a problem for his own endorsement of Lagrangian mechanics outside the hyper-regular domain?
-
-He interprets this as capturing the fundamental kinematical constraints relating position and velocity in classical mechanics.} Central to Curiel's argument is the claim that Lagrangian mechanics naturally encodes the kinematical structure of classical systems whereas Hamiltonian mechanics requires imposing an \textit{ad hoc} condition.
-We will now argue that the condition Curiel claims is ad hoc actually has a natural physical interpretation.
-Additionally, we consider a further disagreement about the proper definition of classical systems, which bears on which formal models are interpreted as physically significant.
-% Additionally, we provide reasons to prefer our definition of classical systems over Curiel's, by considering classical systems outside the hyper-regular domain.
-
-Similar to both us and North, one of Curiel's main goals is to identify the necessary and sufficient mathematical structures for formulating classical mechanics.
-Curiel argues that ``the kinematically-possible evolutions" of a classical system correspond exactly to ``the possible solutions to the Euler--Lagrange equation posed on the tangent bundle" of the system's configuration space \parencites*[292]{Curiel}.
-Hence, he sets out to establish the ``geometrical structure necessary (and manifestly sufficient) for the formulation of the Euler--Lagrange equation having those [kinematically-possible] vector fields as its family of solutions" \parencites*[292]{Curiel}.
-Ultimately, Curiel argues that Lagrangian mechanics respects the necessary and sufficient kinematic constraints that any classical system must satisfy \parencites*[307-308]{Curiel}.
-In contrast, Curiel argues that Hamiltonian mechanics does not inherently respect these kinematical constraints.
-Essentially,
-
-%---interpreted as applying to those systems who possess a unique solution to the Euler--Lagrange equation---. But, Curiel is comfortable talking about Lagrangian systems that are not hyper-regular.
-
-Instead, to treat classical systems according to Hamiltonian mechanics, Curiel argues that we necessarily have to focus on a proper subset of all possible Hamiltonians.
-Nevertheless, Curiel notes that we can formally apply Hamiltonians that do not have the form of Equation~\ref{6.2}, e.g.
-the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$.
-Curiel then argues that all of the ``formally acceptable Hamiltonians" that do not satisfy the quadratic form (\ref{6.2}) are neither physically significant nor physically meaningful.
-It is in this sense that Curiel believes we must artificially restrict Hamiltonian mechanics to those Hamiltonians that satisfy Equation~\ref{6.2}, in order to model physically-meaningful classical systems.
-
-In response, we will defend two points.
-First, we have in fact already in Section~\ref{motion} seen a natural interpretation of both the linear relationship between momentum and velocity and the quadratic Hamiltonian form of equation~\ref{6.2}.
-Both follow from imposing the motion equivalence condition.
-
-% % Another argument one could develop against Curiel: when you move from Newtonian mechanics to a Lagrangian mechanics you appear to lose the ability to describe certain systems.
-A Lagrangian system is fully specified by L(x, v), whereas the Newtonian forces are specified by acceleration as a function of x and v (yieleding three force functions).
-The map between L(x, v) and these forces is a continuous map in the hyper-regular domain, so that a small change in L results in a small change in forces.
-But, outside the hyper-regular domain, there is no unique map from L to these three Newtonian force functions, because there is no single, unique solution.
-So we can write the acceleration as a function of (x, v).
-Whereas in the Newtonian case, we always have a single solution.
-So in other words, the connection between Curiel's equation 4.3 (kinematical constraints defined in terms of the almost tangent structure) vs.
-3.2/3.1 (kinematical constraints in Newtonian mech) is not clear.
-He appears to lose dissipative forces, despite an alleged desire to have Lagrangian mechanics apply to dissipative systems.
-[discussed on 2-1, around 11:25AM]
-
-More broadly, it seems that Curiel simply privileges a different definition of `classical system' than the one we adopt here.
-His definition takes satisfying certain kinematical constraints as fundamental.
-[the state is defined at equal time; the velocity is not] In contrast, we privilege a system having a unique solution, i.e.
-being deterministic and reversible.
-
-% % Another reason to privilege our definition/framework for classical system: states should be defined at equal times, i.e.
-at time slices.
-But velocity is not defined at time slices.
-Velocity is defined on a neighborhood.
-So this is a reason to not view velocity as a state variable, pace Curiel.
-It is a reason to prefer phase space when it comes to specifying the state of the system.
-% It is interesting that conjugate momentum but not velocity is defined at a time slice. Insofar as this is true, shows that they are really not trivial notational variants of each other.
-
-% Incorporated the following into a footnote:
-When it comes to analyzing non-hyper-regular systems, there are two cases.
-In the simplest case, the system simply fails to be hyper-regular at certain points.
-We can thereby decompose the system into hyper-regular subsystems, each with its own Lagrangian (or Hamiltonian).
-The non-hyper-regular points form a boundary that we disregard.
-In the more complicated case, the system fails to be hyper-regular \textit{everywhere}.
-It is difficult to interpret a Lagrangian treatment of such systems, since the Euler--Lagrange equations will not have a single, unique solution.
-
-\end{comment}
-
-%Although we have given reasons to prefer our definition of a classical system, these considerations may not be decisive.
-Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections and summarized in the conclusion.
+% % %, see \textcites{AoPPhy1}. [omit for review].
 
@@
 
-% As additional reasons in favor of Hamiltonian mechanics, 
 Besides this, our approach has many additional interpretive payoffs.
@@
 Whereas the state variables $(q, p)$ suffer from the gauge-dependency of the conjugate momentum, the density provides an invariant description within the Hamiltonian formalism.
-The importance of the density shows that a quantity can be fundamental yet coordinate-dependent.
+Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain.
+Our construction also makes the gauge-dependency of the Lagrangian manifest.
+Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but classical systems in general.
+%The importance of the density shows that a quantity can be fundamental yet coordinate-dependent.
 We have also argued that non-fundamental quantities can be physically significant despite being neither coordinate- nor gauge-invariant.
-Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain while making the gauge-dependency of the Lagrangian manifest.
-Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but classical systems in general.
+[had to cut out since based on OG section 5.2]
 %Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain.
@@
 that mass of the system is constant).
-[Discussed on 2-1, around 11:30AM].
-
-%  It is not our aim here to enter into a burden of proof dispute regarding which definition of `classical system' is to be strictly preferred and why.
-
-% Whereas we happily welcome at least some non-hyper-regular systems into the fold, Curiel seemingly wants to exclude these from the honorific `classical system' \parencites*[305]{Curiel}.
-Hence, Curiel views it as a disadvantage of Hamiltonian mechanics that it applies outside the hyper-regular regime, whereas we view this as a conceptual advantage.
-Regardless of which definition of `classical system' is preferred, Hamiltonian mechanics maintains the mathematical, epistemic, and conceptual advantages described in preceding sections.
-% [To do: but is the argument at the bottom of Curiel page 305 sufficiently compelling?
-Argument seems to appeal to dimensional analysis.
-He argues that Hamiltonian systems outside of this regime include ones that do not make physical sense and violate dimensional analysis.]
-
-%systems whose Hamiltonian takes the form (equation 6.2) $H =\alpha^{m n} p_m p_n + U (q_j) $.
-Here, the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q $.
-
-\begin{comment}
-
-Previous material and notes on Curiel, including additional possible responses:
-
-\begin{quotation}
-Hamiltonian mechanics represents abstract classical systems only in so far as we restrict ourselves to a subfamily of all the formally acceptable Hamiltonians by the \textit{ad hoc} use of conditions foreign to Hamiltonian mechanics itself.
-Its structures do not provide the appropriate concepts and tools to \dots provide any natural justification for the restriction to Hamiltonians of the form (6.2).
-\parencites[306]{Curiel}
-\end{quotation}
-%dots part: `formulate in their terms the required structures that treat configurative quantities differently from momental, nor do they'
-
-% [Set this up properly: Curiel's definition of abstract classical system; Curiel's privileging of certain kinematical constraints, his equations 4.3; Hamiltonian form of equation 6.2; see below].
-Curiel restricts Hamiltonian mechanics to systems whose Hamiltonian takes the form (equation 6.2) $H =\alpha^{m n} p_m p_n + U (q_j) $.
-Here, the functions $U$ and $\alpha^{m n}$ are arbitrary functions of the generalized position coordinates $q $.
-The following provides a helpful summary of Curiel's argument against privileging Hamiltonian mechanics:
-
-We agree with Curiel that his equation 6.2 requires a condition that is imposed by hand, namely the motion equivalence assumption [explain this].
-However, this condition has a clear physical meaning, and it corresponds to a physical condition that must hold in order to define/apply Lagrangian mechanics.
-\textit{Pace} Curiel, there exist Hamiltonian systems that do not satisfy his equation 6.2 (equivalently, classical systems where the momentum is not a linear function of the velocity), and these will even have a Lagrangian, although they will not be an abstract classical system in Curiel's sense.
-Hence, there seems to be a burden on Curiel to distinguish---in a principled way---the classical Lagrangian systems from those that are not classical.
-% Hence, we agree with Curiel that ``it is not necessary that [these conditions, Eqn 6.2] hold in order for the theory to be meaningfully applied to model a [classical] system" [305].
-Whereas Curiel takes this as an argument for the primacy of Lagrangian mechanics, we take it as an argument for the primacy of Hamiltonian mechanics, since Curiel's conditions are equivalent to imposing motion equivalence.
-
-% Curiel's main point: the representation of a classical mechanical system in the Lagrangian framework is \textit{natural} in a way that its representation in a Hamiltonian formulation is not.
-The former representation depends on intrinsic structures of the classical mechanical system rather than an artificial construction (270).
-
-Rebuttal 2: On its own, the motion of a system in physical space is not sufficient for defining the system's dynamics.
+
+%On its own, the motion of a system in physical space is not sufficient for defining the system's dynamics.
 The kinematics/motion does not suffice to determine the forces acting on the system.
@@
 These two distinct classical systems can nevertheless have the same equations of motion.
-% The Hamiltonian framework does not face a corresponding underdetermination problem because it does not allow for dissipative systems.
+% The Hamiltonian framework does not face a corresponding underdetermination problem because it does not allow for dissipative systems [at least when we restrict to deterministic/reversible evolN].
 Hence, there is a sense in which the Lagrangian formalism is more mathematically general than the Hamiltonian formalism.
@@
 
-Rebuttal 3: The standard predator--prey model from differential equations satisfies Curiel's definition of an abstract classical system.
-However, it does not satisfy Curiel's notion of the split between configurational vs.
-velocital coordinates.
-And it is not true that the velocity will be the first derivative of the position.
-Moreover, the predator--prey system is neither Hamiltonian nor Lagrangian, because it is not necessarily deterministic and reversible.
-% Related to equation 3.1. Make precise.
-% Furthermore, the momentum does not need to be a linear function of velocity. For example, the photon does not satisfy this relationship.
-
-Rebuttal 4: make precise and substantiate with textual evidence (might be relevant for responding to objection considered below): implicitly, Curiel assumes that classical systems are defined as Newtonian systems in an inertial frame with Cartesian coordinates.
-Implicitly, Curiel assumes that energy is a function of position and velocity, entailing that we have a Hamiltonian.
-The claim in footnote 17 that the kinetic energy is a function of velocity depends on the choice of a particular coordinate system; it is not a coordinate-invariant claim.
-Curiel's thought experiment assumes that we have a constant force (in this case gravity) and that we are in a reference frame where we can compare forces (e.g.
-an inertial frame).
-
-[Most sophisticated objection I can think of on behalf of Curiel:] Like our argument, part of Curiel's argument involves appealing to systems outside of the hyper-regular domain.
-Curiel acknowledges both that only hyper-regular Lagrangians have unique solutions and that Hamiltonian systems have unique solutions even outside of the hyper-regular domain \parencites*[291, 308]{Curiel}.
-Nevertheless, he argues that this is actually an advantage of Lagrangian mechanics, rather than a disadvantage.
-On his view, the unique solutions of Hamiltonian mechanics outside the hyper-regular domain is precisely why ``one must impose the quadratic form of the Hamiltonian as an \textit{ad hoc} condition" \parencites*[308]{Curiel}.
-In contrast, he argues that Lagrangian mechanics precisely delimits the proper domain of abstract classical systems, namely those that satisfy a particular set of physically significant kinematical constraints (his equations 4.3).
-Although Lagrangian systems do not admit unique solutions outside the hyper-regular domain, these non-unique solutions will still satisfy the kinematical constraints of Lagrangian mechanics \parencites*[311]{Curiel}.
-
-Figure out how to relate the following to the preceding: In a footnote, Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
-There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other" \parencites*[1179]{Barrett2}.
-Indeed, outside the hyper-regular domain, the Euler--Lagrange equations do not have a single, unique solution.
-Hence, it is difficult to say what a Lagrangian system even is outside the hyper-regular domain.
-As an example, consider the case where the Lagrangian $L$ equals zero.
-Then for this ``system," all paths have the same action, so all paths are equally possible.\footnote{This Lagrangian corresponds to a Hamiltonian that also equals zero, entailing that $dH = 0 $.
-By Hamilton's equations, the derivatives of the position and momenta are zero, indicating that this system is static.} In contrast,  a Hamiltonian defines a unique solution (unique evolution in phase space) even when outside the hyper-regular domain (i.e.
-when the Hessian is zero).
-
-Summary: Hamiltonian mechanics provides solutions outside hyper-regular domain (but not these kinematical constraints) whereas Lagrangian mechanics preserves these kinematical constraints outside hyper-regular domain.
-Worry: perhaps GC and Curiel are in a verbal dispute about the proper/privileged definition of a classical system (with Curiel privileging these kinematical constraints, and GC privileging deterministic/reversible evolution).
-
-On pages 309-310, Curiel considers a Hamiltonian that could model the photon (so it seems).
-In considering a Hamiltonian system that does not respect the Lagrangian kinematical constraints, Curiel considers the Hamiltonian $H=p$, which corresponds to a system moving with constant velocity (e.g.
-velocity of 1).
-So this could very well be a model of a photon, moving at the speed of light.
-Curiel points out that the corresponding Lagrangian: $L= vp - H = \dot{q}-1$, yields a trivial Lagrangian problem: $0=0$ upon evaluating the Euler--Lagrange equation.
-Upshot: ``not all Hamiltonian vector fields map to second-order vector fields under the inverse-Legendre transform" [310].
-But then does Curiel consider this system to not be a classical system?
-
-[To do: figure out if the following claims made by Curiel regarding the differences between the symplectic structure in Hamiltonian mechanics (which is fixed) vs.
-the symplectic structure that arises from the Euler--Lagrange equations (which varies from Lagrangian to Lagrangian) are relevant to our argument or rebuttals: \\
-Hamiltonian mechanics: ``in Hamiltonian mechanics, the fact that the symplectic structure is fixed once and for all enforces the satisfaction of the kinematical constraints (5.3), and it encodes the existence and uniqueness of solutions to the equations of motion and the conservation of energy, among many other things.
-All of these features are independent of the individual Hamiltonian used in the equation" [310].
-\\
-Lagrangian mechanics: ``In Lagrangian mechanics, in contrast, it is the Lagrangian itself that encodes existence and uniqueness of solutions (which will obtain when and only when the Lagrangian is regular), and there is not even a guarantee that energy is conserved, as one can, for example, represent dissipative systems in the Lagrangian framework" [310].
-
-% Either way, presumably we can agree with Curiel's explanation/understanding for why the Lagrangian must be regular: Curiel arguing that he can give an account of ``why it makes physical sense to demand that the Lagrangian be regular", i.e.
-satisfy equation 4.6 for the Hessian determinant.
-He notes that by considering the variational problem, we understand ``why the Lagrangian must be quadratic (at least) in the velocities for there to be a unique solution to the Euler--Lagrange equation" [308].
-Argument: for the variational problem to be well-posed, you need to be able to ``take non-trivial derivatives of the Lagrangian up to second-order in the velocity, since the crucial integration by parts that yields the Euler--Lagrange equation demands this, on pain of giving, for example, the tautology $0=0$" in some cases, or a contradiction in others.
-
-\end{comment}
-
-\begin{comment}
-
-\section{Material looking for a possible home}
-
-[To do: not sure if following warrants discussion.
-Could possibly connect this with the earlier discussion of the significance of defining the unit system, since that seems key to our response to this hypothetical worry.]
-Teh and Tsementzis \parencites*[]{Teh} provide another argument for the equivalence of Lagrangian and Hamiltonian mechanics within the hyper-regular domain.
-They develop this using Tulczyjew's reformulation of Hamiltonian mechanics, which shows that ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" \parencites*[46]{Teh} [I slipped in part of this claim as a footnote into section 2.1: shows another reason to view Hamiltonian mechanics as fundamental in a non-metaphysical sense].
-Rest of the discussion seems too complicated and technically involved to set up properly.
-Response-claims below might not even be completely accurate.
-Again, there is a worry that this equivalence result might undercut our privileging of Hamiltonian mechanics.
-However, the equivalence is again just within the hyper-regular domain, requiring that motion equivalence is satisfied.
-Furthermore, Tulczyjew's formulation relies heavily on the notion of a Lagrangian submanifold, since ``a mechanical system turns out to be a Lagrangian submanifold of an appropriately defined space" \parencites[47]{Teh}.\footnote{A submanifold $N$ of a symplectic manifold $(P, \omega) $ is Lagrangian provided that i) it has half the dimension of P and ii) the pullback of the symplectic form $\omega$ vanishes along the inclusion of N in P \parencites[48]{Teh}.} Yet importantly, Lagrangian submanifolds do not distinguish between the space of velocities and the space of momenta, i.e.
-they obscure the important physical differences between these two kinds of quantities.
-Not every constructible Lagrangian submanifold defines units in a physically significant way.\footnote{One could even construct a suitable Lagrangian submanifold by mixing position and momenta coordinates.} It is the physical significance of defining a unit system that breaks the formal (syntactic) symmetry between position and momentum coordinates.
-% [To do: verify the veracity of this discussion of Lagrangian sub manifolds]
-%Furthermore, within this reformulation, Lagrangian submanifolds play a very important role. Does this lead to a possible privileging of Lagrangian mechanics?
-
-[To do: Not sure where to put the following material, if anywhere; perhaps after a discussion of why we restrict to hyper-regular domain?:]
-
-Notes/questions from 11-16-20 meeting
-\begin{enumerate}[1.)]
-
-\item Possible connection with motion equivalence principle: Sameness of empirical content (empirical equivalence) is cashed out in terms of agreeing ``on the allowable trajectories for particles in the system" [1180].
-For a Lagrangian system, this is characterized in terms of a base integral curve of the Lagrangian vector field $(X_L)^a$ that ``encodes one way that the positions of particles in a system with Lagrangian L might evolve over time", while the base integral curves of the Hamiltonian vector field perform an analogous role [1180]
-
-\begin{itemize}
-
-\item   Marsden and Abraham 1978, Theorem 3.6.2 proves that (in the hyper-regular domain) the Lagrangian vector field and the corresponding Hamiltonian vector field (obtained via Legendre transformation) have the same base integral curves.
-This means that they ``agree on how the positions of particles will evolve over time" [1181].
-Is this the same as the motion equivalence assumption?
-
-\end{itemize}
-
-\item Barrett footnote 18 wonders whether there could be a ``correspondingly general formulation of Lagrangian mechanics" that is analogous to the symplectic but-not-cotangent-bundle formulation of Hamiltonian mechanics.
-Does GC derivation show that there cannot be such a correspondingly general formulation of Lagrangian mechanics?
-
-\begin{itemize}
-
-\item: GC in response: defining the Lagrangian presupposes that we have trajectories of positions over time.
-Velocity is then a derived quantity of these positions.
-So the tangent bundle seems to be intimately connected with having trajectories.
-Hence, if we eliminated reference to the tangent bundle, we would seemingly not be able to define trajectories.
-
-\item Furthermore, there is a sense in which GC derivation recovers the tangent bundle and shows why it must be well-defined: provided the evolutions are differentiable (which is needed to preserve densities), it follows that velocities are well-defined, and hence so is the tangent bundle.
-Note that the trajectories alone do not entail that velocities are well-defined, since the trajectories could in principle not be differentiable.
-
-\item Note that although the symplectic manifold already has some global structure (which corresponds to the global tangent bundle), the cotangent bundle itself is in some sense more global (and hence more specialized).
-Helpful analogy: a manifold can be locally Euclidean (which is a global future) without being globally Euclidean.
-The sense in which the tangent bundle and symplectic manifold structure are global is analogous to being locally Euclidean, whereas the global cotangent bundle structure is analogous to being globally Euclidean.
-So in going from the cotangent formulation to the symplectic formulation, we are becoming less specific (more general).
-To generalize the Lagrangian, we would seemingly need to lose some structure, but the action seems to be global (integral defined over a trajectory).
-
-\item It might be worth noting these differences, as a way of motivating the claim that there probably is not a correspondingly general formulation of Lagrangian mechanics.
-
-\end{itemize}
-
-\item Barrett also shows that the vector field formulations of Lagrangian and Hamiltonian mechanics are not categorically equivalent [1185].
-He interprets this as substantiating Curiel's claim that the ``Legendre transform does not respect the kinematical constraints of Lagrangian mechanics", namely because it shows that ``there are Lagrangians that give rise to the same Lagrangian vector field, but do not translate---via the Legendre transformation---into Hamiltonians that give rise to the same Hamiltonian vector field" [1186].
+%Barrett also shows that the vector field formulations of Lagrangian and Hamiltonian mechanics are not categorically equivalent [1185].
+He interprets this as substantiating Curiel's claim that the ``Legendre transform does not respect the kinematical constraints of Lagrangian mechanics'', namely because it shows that ``there are Lagrangians that give rise to the same Lagrangian vector field, but do not translate---via the Legendre transformation---into Hamiltonians that give rise to the same Hamiltonian vector field'' [1186].
 What does this mean in the context of motion equivalence?
-
-\begin{itemize}
-
-\item Barrett interprets this situation as simply showing that the Hamiltonian vector field contains more information about the Hamiltonian than the Lagrangian vector field does about the Lagrangian.
+%Barrett interprets this situation as simply showing that the Hamiltonian vector field contains more information about the Hamiltonian than the Lagrangian vector field does about the Lagrangian.
 If two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to a arbitrary constant.
@@
 
-\end{itemize}
-
-\item Teh and Tsementzis discuss Tulczyjew's reformulation of Hamiltonian mechanics, which shows that ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry" [PDF 3].
-Does this accomplish a similar moral as the assumptions of physics framework?
-
-\item Furthermore, within this reformulation, Lagrangian submanifolds play a very important role. Does this lead to a possible privileging of Lagrangian mechanics?
-
-\end{enumerate}
-
-\end{comment}
-
-\printbibliography
+\bs
+\printbibliography[title ={References}]
+\es
 

### 2021-04-21 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: Andrea's review of sec 4

@@
 We assume that, after time $t$, the hatch is opened and that at time $\hat{t}$ the ball has settled.
-As before, $\hat{q} \vee \hat{r} \equiv \top$ because the ball must be in one of the two positions after it dropped, therefore, if $p$ is true, the left side of the biconditional is true while the right side is false.
+As before, $\hat{q} \vee \hat{r} \equiv \top$ because the ball must be in one of the two positions after it dropped, therefore, if $p$ is true, the left side of the bi-conditional is true while the right side is false.
 
@@
 
-\section{Math}
+\section{On the Common Logical Structure between Quantum and Classical Mechanics}
 \label{Math}
 
-In this section we present our main argument, which is that the state space of quantum mechanics and the space of classical probability distributions possess a similar structure that allows one to apply distributive propositional calculus (i.e.
-``classical logic'') on the first and non-distributive propositional calculus (i.e.
-``quantum logic'') on the second.
-That is, both theories will have two lattices of propositions: a distributive one that follows classical logic and a non-distributive one that follows quantum logic.
+After having critically assessed the general motivations to introduce quantum logic as a consequence of the physical content of QM, in this section we present our main argument according to which it is possible to assign a distributive lattice to quantum propositions.
+Remarkably, in what follows we will see that the non-distributivity of the quantum logic lattice does neither originate from non-commuting observables nor from superpositions of quantum mechanics: it is a property of the lattice of subspaces of \emph{any} vector space and of the lattice of subgroups of any group more in general (cf.\ \cite{Davey:2002}).
+To this regard, indeed, we also show that taking into account the space of classical probability distributions one can apply a non-distributive lattice to classical propositions.
+This fact is due to the similar structures of the state space of quantum mechanics and the space of classical probability distributions.
 
-Note that the classical space we will be using is not phase-space itself, but the space of probability distributions over phase-space, which is a real Hilbert space and thus has a similar structure to the quantum counterpart.
-The central point is that many features of quantum logic stem not from quantum mechanics per-se, but from the fact that quantum mechanics is a probabilistic theory, which means a classical probabilistic theory will share the same features.
-We find that the main point of confusion stems from comparing quantum theory, which is inherently a probabilistic theory, to a non-probabilistic classical theory.
-This approach does not allow one to understand which differences are due to the classical/quantum nature of the theory, and which are due simply to the probabilistic/non-probabilistic nature.
-By comparing quantum theory to a probabilistic classical theory, we put ourselves in a position to fully understand which difference is due to what, and have a much clearer understanding of both.\footnote{Note that probabilistic theories are important even in classical mechanics.
-Due to the nature of experimental verification, even in classical mechanics, there are always going to be uncertainties in the preparation and measurements, and repeated measurements will always lead to statistical constructs.} Critically, we will see that the non-distributivity of the quantum logic lattice does not originate from non-commuting observables or superposition of quantum mechanics: it is a property of the lattice of subspaces of any vector space and, indeed, of the lattice of sugroups of any group more in general.
-[CITE: Introduction to Lattices and Order Theory, Davey, Priestley - 2002 CUP]
+In what follows, then, it will be argued that both theories---quantum and classical---may be assigned \emph{two} lattices of propositions: a distributive one that follows the laws of classical logic, and a non-distributive one obeying the rules of quantum logic.
+Contrary to the widespread opinion, therefore, we will claim not only that quantum physics does not necessarily entails a revision of classical propositional calculus, but also that it is actually possible to show that quantum propositions can form a distributive lattice.
+Furthermore, in agreement with the example of the violation of the distributivity law in the classical context seen in the previous section, in the remainder of the section we will argue also that the propositions of classical mechanics may obey a non-classical logic.
 
-Mathematically, the distributive lattice will be provided in both classical and quantum mechanics by the Borel algebra, the smallest $\sigma$-algebra that contains the topology; the non-distributive lattice will be provided by the set of subgroups ordered by inclusion.
-We will see that the non-distributive lattice is a strict subset of the distributive one, meaning that everything that can be described in quantum logic can also be described in classical logic.
-Moreover, the distributive lattice will include propositions, like propositions on the expectation value of observables, that are of scientific interest.\footnote{In an experimental context, in fact, by ``measurement'' one typically refers to the statistical collection of many ``single takes''.
-Quantum logic describes the single-take measurements.
-For example, measurements of cross sections, the most typical type of measurement in the context of particle physics, are not part of the quantum lattice, and neither are measurements of the probabilities themselves.}
+Before we proceed, it is worth noting that the classical space we will be using is not phase-space itself, but the space of probability distributions over phase-space, which is a real Hilbert space and thus has a similar structure to its quantum counterpart.
+To this regard, it should be stressed that probabilistic theories are important even in classical mechanics.
+Due to the nature of experimental verification, also in the context of classical mechanics there are always going to be uncertainties in the preparation and measurements, and repeated measurements will directly lead to statistical constructs.
+However, we want to emphasize that this paper aims primarily at showing that QM can be given a distributive lattice of propositions.
+Here we want to enter neither into debates concerning the ontology of quantum mechanics, nor into discussions concerning the similarities between the commitments of QM and those of classical mechanics.
+In this essay we simply point out a formal similarity between the above mentioned mathematical structures that has implications for the logic of both theories.
 
-%they do not seem well-known among physicists and philosophers of science, therefore we briefly summarize their definition, use and meaning.
-[CITE: Kelly, Logic of Reliable Inquiry - Carcassi/Aidala, Assumptions of Physics]
+From a mathematical perspective, the distributive lattice will be provided in both classical and quantum case by a Borel algebra, which is the smallest $\sigma$-algebra on a set $X$ equipped with a topological structure that contains the topology; the non-distributive lattice will be provided by the set of subgroups ordered by inclusion.
+We will claim that the non-distributive lattice is a strict subset of the distributive one; this result entails that everything that can be described in quantum logic can also be described in classical logic.
+Moreover, the distributive lattice will include propositions, like sentences concerning the expectation value of observables that are of high scientific interest, since in an experimental context by ``measurement'' one typically refers to the statistical collection of many ``single takes''.
+On the contrary, it is well-known that standard quantum logic deals only with single-take measurements, remaining silent w.r.t.\ propositions involving statistical statements.
+Hence, measurements of cross sections, the most typical type of measurement in the context of particle physics, are not part of the quantum lattice, and neither are measurements of the probabilities themselves.
+Then, to show that one may apply a distributive lattice to quantum propositions can be seen as an important result in order to extend the set of meaningful logical propositions of quantum mechanics.
+To provide an extended quantum logic lies beyond the introductory scope of the present paper and will be the subject for future research; here we limit ourselves to show that such an extension is possible.
 
-%It is useful, at this point, to review $\sigma$-algebras, which are foundational structures in mathematics,
+\subsection{The Distributive Lattice of Quantum Mechanics}
 
-Note that mathematical branches important in the context of physics, like measure theory [CITE: Measure theory, by Donald L.
-Cohn] and probability theory,[CITE: Grimmet/Stirzaker, Probability and Random Processes] are built not top of $\sigma$-algebras, which should be considered foundational structure.
-Let us therefore review some crucial aspects.
+As anticipated above, in this section our arguments are fundamentally based on the notion of $\sigma$-algebra, which is a significant structure in the context of foundational research, given that physically relevant mathematical branches like measure theory (cf.\ \cite{Cohn:2013}) and probability theory (cf.\ \cite{Grimmett:2001}) are built on top of it.
+Let us then review some of its essential features.
 
-We start with a set $X$.
-In the quantum case, this will correspond to the set of all possible quantum states.
-In the classical case, this will corrsepond to the set of all possible probability distributions over phase space.
-We should stress, however, that $\sigma$-algebras are defined independently of what the elements of $X$ actually are.
-For a simpler example, $X$  could correspond to the set of the real numbers for a continuous quantity.
-A $\sigma$-algebras is a collection of subsets of $X$, closed under countable union, countable interesection and complement with respect to $X$, meaning that performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra.
-For example, a standard $\sigma$-algebras for the real numbers will typically contain all the open intervals $(a, b)$.
+In order to begin our discussion, let us consider in the first place a set $X$: in the quantum case it will correspond to the set of all possible quantum states; similarly, in the classical case it will correspond to the set of all possible probability distributions over phase space.\footnote{We should stress, however, that $\sigma$-algebras are defined independently of what the elements of $X$ actually are.
+For a simpler example, $X$  could correspond to the set of the real numbers for a continuous quantity.} By definition, a $\sigma$-algebra is a collection of subsets of $X$ closed under countable union, countable intersection and complement with respect to $X$; the closure condition requires that  performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra.
+Interestingly, these sets represent statements concerning the objects described by $X$, i.e.\ quantum states and classical probability distributions.\footnote{In the quantum case this is exemplified also in Section \ref{QL}.} Then, in virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets.
+Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement.
+Hence, the $\sigma$-algebra generates a Boolean algebra of statements.\footnote{Note that the interpretation used here is exactly the same used in probability theory.
+In fact, a probability space is defined by three elements: a sample space $\Omega$ which represents the set of all possible outcomes; a $\sigma$-algebra over $\Omega$ which represents all events, all propositions; and a measure that assigns a probability to each event.}
 
-A collection of subsets may seem quite an abstract concept, but it can be readily made concrete: any statement on the objects represented by $X$ can be represented by a set.
-For example, the open set $(a, b)$ will exactly correspond to the statement ``the value of the quantity is between $a$ and $b$''.
-It is therefore useful to think about $\sigma$-algebras as collections of statements about $X$ instead of sets.
-As the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement: the $\sigma$-algebra can be thought as a Boolean algebra of statements.
-Note that this interpretation is exactly the one used in probability.
-In fact, a probability space is defined by three things: a sample space $\Omega$ which represents the set of all possible outcomes; a $\sigma$-algebra over $\Omega$ which represents all events, all propositions; and a measure that assigns a probability to each event.
-
-Every topological space comes already equipped with a $\sigma$-algebra called Borel algebra.
-A topology, in fact, is also a collections of sets, called open sets, and the Borel algebra is defined to be the smallest $\sigma$-algebra that contains all the open sets.
-The topology will not be directly important for us, but only as a mean to construct the Borel algebra.
-Complete normed spaces, such as Hilbert spaces, have a canonical way to construct a topology and therefore a $\sigma$-algebra.
-Let us see first how it works in a standard three-dimensional Euclidean space.
+In the second place, it is crucial to underline that every topological space---as for instance the complex Hilbert space of quantum mechanics and the real Hilbert space of classical probability distributions considered in this essay---comes already equipped with a $\sigma$-algebra called Borel algebra.\ A topology, in fact, is also a collection of sets---called \emph{open sets}---and the Borel algebra is defined to be the smallest $\sigma$-algebra that contains all the open sets.\footnote{Here the topology is important for us only as a mean to construct the Borel algebra.} Complete normed spaces, such as the mentioned Hilbert spaces, have a canonical way to construct a topology and, therefore, a $\sigma$-algebra.
+Let us see how this procedure works in the easier case of a standard three-dimensional Euclidean space.
 Given two points $x$ and $y$, we can write the square of distance $\epsilon^2 = |y - x|^2$ as the square of the norm of the vector between the two points.
@@
 These are exactly the open sets that generate the topology, and the Borel algebra will be the smallest $\sigma$-algebra that contains these sets.
-This works in the same way for all normed spaces.
+This works analogously for all normed spaces.
 A Hilbert space, in particular, comes with an inner product which also defines a norm.
 We can define a distance $\epsilon^2 = |\psi - \phi|^2=\langle \psi - \phi , \psi - \phi \rangle$ between two states, and the statement ``the state of the system is within $\epsilon$ of $\psi$'' corresponds to the set $U = \{\phi \; | \; |\psi - \phi|^2 < \epsilon^2\}$.
-As we'll see later, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.\footnote{The choice of the square root of the density over the density itself is really a technical choice.
-Of all the $L^p$ spaces, the only Hilbert space is $L^2$, the space of the square-integrable functions.} We will have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
-In all normed spaces, the topology is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.\footnote{In the case of Hilbert spaces and quantum mechanics, the map between topology and experimental verification does not work perfectly because absolute modulus and phase are not physically meaningful.
-This can be fixed by using the corresponding projective space, which removes the overdetermination.
-We are not going to do this as projective spaces have their own peculiarities and are used less.}
+As we have already stated, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.\footnote{The choice of the square root of the density over the density itself is merely a technical choice.} In this case we have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
+In all normed spaces, the topology is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.
 
-It is a simple result that the Borel algebra thus constructed contains all propositions of quantum logic.
+It is a simple result that the Borel algebra thus constructed contains all propositions of quantum logic, and we state it in the following proposition:
 
@@
 Let $p \in L(H)$ be a closed sub-space of $H$.
-Then $p$ is a closed set, the complement of an open set, and therefore a Borel sets: $p \in \Sigma$.
+Then $p$ is a closed set, the complement of an open set, and therefore a Borel set: $p \in \Sigma$.
 Let $v \in H$ be a vector in the Hilbert space and consider the singleton $U = \{ v \}$.
 This is not a closed sub-space and therefore $U \notin L(H)$.
-As the topology of a Hilbert space is Hausdorff, all singletons are closed and therefore $U \in \Sigma$.
+\textcolor{red}{As the topology of a Hilbert space is Hausdorff, all singletons are closed and therefore $U \in \Sigma$.
 This means $L(H) \subset \Sigma$.
-This concludes the proof.
+}This concludes the proof.
 
-In quantum logic, the propositions corresponded to closed subspaces, which are closed sets and are therefore Borel sets.
-All the propositions of quantum logic are therefore readily found in the Borel algebra of the Hilbert space.
-The Borel algebra, which is a Boolean algebra that follows classical logic, contains also all the statements of quantum logic.
-This construction also applies identically for the classical cases, where we will also be able to talk about the set of all closed subspaces which will have a parallel structure.
+The above proposition contains our first crucial conclusion.
+As we stated in Section \ref{QL}, in quantum logic the propositions correspond to closed subspaces, which are closed sets and, thus, are Borel sets.
+All the propositions of quantum logic are consequently readily found in the Borel algebra of the Hilbert space.
+The Borel algebra, in turn, is a Boolean algebra that follows classical logic, and contains all the statements of quantum logic.
+This construction also applies identically for the classical cases, where we will also be able to talk about the set of all closed subspaces.
 
-While all of this may seem daunting to someone who has never seen this material, we want to stress that these are very standard constructions.
-We are not taking some obscure framework from some fringe field of mathematics: these are the standard mathematical techniques used to study these spaces and can be found in any textbook.
-[CITE: Elements of Hilbert Spaces and Operator Theory, Harkrishan Lal Vasudeva, https://www.springer.com/gp/book/9789811030192 ] What we are doing here is bringing out their full physical significance, which is too often overlooked.
+While all of this may seem suspect to some reader, we would like to stress that these are very standard constructions.
+Alternatively stated, all of what has been mentioned so far are the standard mathematical techniques used to study these spaces and can be found in any textbook (cf.\ \cite{Rudin:1991}, \cite{Vasudeva:2017}).
+Here we are bringing out the full physical significance of these mathematical constructs, which is too often overlooked.
 Understanding the link between the mathematical structures and the physics they represent is crucial to clearing out possible misunderstandings.
@@
 
-One may ask: do these extra sets correspond to physically meaningful propositions? To answer that, let us first demonstrate the following proposition.
+Arrived at this point, one may ask: do these extra sets correspond to physically meaningful propositions? To answer this question, let us first demonstrate the following proposition.
 
@@
 Borel sets also allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
-This is NOT equivalent to the statement ``the particle is in the eigenstate $x_0$''.
-The second will be true only if the whole function is at $x_0$.
+This is \emph{not} equivalent to the statement ``the particle is in the eigenstate $x_0$''.
+This latter proposition will be true only if the whole function is at $x_0$.
 That is, if the $x_0$ is the support of the wave-function.
-The first is a lot less stringent: the wave-function can even be distributed over an infinite range (e.g.
-a Gaussian wave packet) just as long the average is $x_0$.
-This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.
-a Gaussian wave packet can be centered on any value of position and momentum).
+The former is a lot less stringent: the wave-function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$.
+This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.\ a Gaussian wave packet can be centered on any value of position and momentum).
 
@@
 The Borel algebra contains as a sub-algebra the statements we need for (at least one form of) the classical limit.
-And this is a purely classical subalgebra, because the average position and average momentum can both be well defined.
+And this is a purely classical sub-algebra, because the average position and average momentum can both be well defined.
 Moreover, note that this is not dissimilar to what happens in classical mechanics when, for example, we study planetary motion using only the position and momentum of the center of mass.
@@
 
-Now that we have shown that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us do the converse: show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
-Given that quantum mechanics is a probabilistic theory, let us compare it to a classical probabilistic theory.
-So let us consider the space of all possible probability distributions $\rho(x,p)$ over phase-space.
-We can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$.
+\subsection{The Non-Distributive Lattice of Classical Mechanics}
+
+Now that we have argued that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
+Given that quantum mechanics is inherently a probabilistic theory, let us compare it to a classical probabilistic theory.
+Thus, let us consider the space of all possible probability distributions $\rho(x,p)$ over phase-space.
+As already stated, we can formalize this as a Hilbert space, where a vector $\psi$ corresponds to $\sqrt{\rho(x,p)}$.
 This will be a real vector space, and we are only going to consider operators that can be expressed as a multiplicative function $f(x,p)$.
@@
 To really understand the source of the non-distributivity, let us study the simplest possible example.
-Let us assume that there are only two possible classical states, $1$ and $2$.
+Let us assume that there are only two possible classical states of a certain physical system $s$, $1$ and $2$.
 A state is identified by a two-valued distribution $[\rho_1, \rho_2]$, where $\rho_1$ is the probability to find the object in state 1 and $\rho_2$ is the probability to find the object in state 2.
 We note $[k,0]$ the subspace spanned by the first state only.
-This corresponds to the statement ``the system is certainly in state $1$'' since the probability for state $2$ is zero.
-Conversely, $[0,k]$ corresponds to the statement ``the system is certainly in state $2$''.
+This corresponds to the statement ``the system $s$ is certainly in state $1$'' since the probability for state $2$ is zero.
+Conversely, $[0,k]$ corresponds to the statement ``the system $s$ is certainly in state $2$''.
 The disjunction $[k,0] \vee [0,k]$ will correspond to the subspace spanned by both, which is the whole space $[k,j]$ where we can assign different non-zero probability to either case.
@@
 It's a property of the lattice of subspaces of any vector space and, more generally, of the lattice of subgroups of any group.
-In fact, consider the group of boost transformations in relativity.
+Indeed, as a further example, consider the group of boost transformations in relativity.
 The boosts along one direction form a subgroup.
@@
 We want to be clear that we are not claiming that the lattice of subspaces of quantum logic is not of interest, or that it is of equal interest in both classical and quantum mechanics.
-We are pointing it out that the argument that quantum mechanics REQUIRES a radical departure from classical logic in untenable, given both theories follow similar structures.
-Such arguments have to be a lot more nuanced.
-Additionally, once we understand the similarity between the theories, we are in a much better position to understand the differences, which naturally are there.
+Here we have pointed it out that the argument that quantum mechanics \emph{requires} necessarily a radical departure from classical logic is untenable, given that both theories are built on similar structures.
+We conclude that such claims should be substantially weakened, in the light of the arguments that we proposed in this section.
 
-For example, quantum mechanics can only be expressed as a probabilistic theory and therefore ALWAYS requires a structure that in classical mechanics is not mandatory.
+Finally, let us conclude by saying that once we understand the similarity between the theories, we are in a much better position to understand their differences, which naturally exist.
+For example, quantum mechanics can only be expressed as a probabilistic theory and therefore \emph{always} requires a structure that in classical mechanics is not mandatory.
 In quantum mechanics, all subspaces of dimension one represent a pure state while in classical mechanics this is not the case.
@@
 
-\section{Discussion}
-\label{Discussion}
-%\section{Old notes}
-%
-%* We can always use the Borel algebra for our statements. This will include(?) all propositions of ``quantum logic''.
-%
-%It includes (?) also proposition on expectation values, say of position and momentum. Those will behave exactly like ``point particles''. It also includes (?) propositions on the support.
-%
-%Show that this is similar to classical distributions.
-%
-%
-%Claim: in quantum mechanics the ``correct'' logical structure the is algebra of subspaces of the Hilbert space.
-In this algebra (which is a $\cap$-structure) the meet functions differently than in Boolean logic.
-Therefore it does not follow classical logic.
-%
-%Resolution: that algebra does not include many statements that are interesting experimentally.
-For example, it does not include statements about the expectation values which, in the end, are the result of experiments.
-(Cross-section?) One
-%
-%
-%Technical view: the state space of quantum states is a separable Hilbert space; it is a Banach space (complete normed space) and admits a norm-induced topology; it is separable, so it admits a countable basis, so the topology is second countable; therefore the state space has a countably generated Borel-algebra.
-This, like any set theoretic structure, follows Boolean algebra.
-Each state is a point and state-statements like ``the system is in state $s_1$'' are singletons in the $\sigma$-algebra.
-State-statements are all incompatible with each other: either we prepared one state or we prepared another.
-%
-%If a process with an initial and final state (the internal are disregarded), we have two lattices of propositions.
-The sample space is the Cartesian product, the $\sigma$-algebra is the product.
-A measurement process will follow this pattern.
-We can say ``the system was prepared in state $s$'' and ``the system was measure in state $s$'' and there are two different statements.
-In the product algebra, we can then mix and match statements with the usual Boolean constructs.
-Both these are analogous in a classical and quantum case.
-These will naturally result in different lattices, but they will still be Boolean lattices.
-%
-%What changes are the probability spaces we can construct.
-In the quantum case, the probability is not in general a probability measure and has to be consistent with the Born rule (i.e.
-the inner product).
-
-\section{Conclusion}
+\section{Discussion and Concluding Remarks}
 \label{conc}

### 2021-04-23 Overviews/Informational granularity/InformationGranularity.tex
InfGran: review with Mark

@@
 
-\begin{prop}
+\begin{prop}\label{fineness_monotony}
 Other properties satisfied by fineness:
@@
 
+\begin{defn}
+	Let $\stmt_1, \stmt_2 \in \tdomain$ be two statements such that $\stmt_1 \finer \stmt_2$.
+We say that they are \textbf{finitely comparable} to each other if we can find a finite set $\{ \stmt[t]_i \}_{i=1}^n$ such that $\stmt[t]_i \finer \stmt_1$ for all $i \leq n$ and $\stmt_2 \finer \bigOR_{i=1}^n \stmt[t]_i$.
+\end{defn}
+
+\begin{prop}
+	If two statements are finitely comparable then they cannot be infinitesimals with respect to each other.
+\end{prop}
+\begin{proof}
+	TODO
+\end{proof}
+
 \begin{prop}\label{infinitesimal_conjunction}
@@
 
+The following property is crucial for the theory of qualitative probability, but fortunately it is immediate from either definition.
+
+\begin{prop}
+For all statements $\stmt$ and $\stmt[u]$, we have $\impossibility\lequ\stmt\lequ\certainty$.
+\end{prop}
+\begin{proof}
+	TODO
+\end{proof}
+
 \begin{prop}
@@
 
-\begin{conj}
+\begin{desid}
 	The relation $\lequ$ is reflexive and transitive and therefore a preorder. 
-\end{conj}
-
-Fortunately, the current definitions of $\lequ$ and $\equ$ behave well together:
+\end{desid}
+
+Fortunately, at least the first definitions of $\lequ$ and $\equ$ behave well together:
 
@@
 
+For the second definition of $\lequ$, transitivity of $\equ$ is needed which we have not proven.
+
 Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotone?
+
+Eventually, we wish to use $\lequ$ as a total pre-order in order to use it as a qualitative probability.
+We will need an interesting-enough subset of $\tdomain$ to capture the desired notions while still having $\lequ$ as a total order.
 
@@
 
-\begin{defn}[Monotone continuity]
+\begin{defn}[Monotone continuity]\label{monotone_continuity}
 	Given a domain $\tdomain$ and a unit $\stmt[u] \in\tdomain$, we say that $\lequ$ satisfies \textbf{monotone continuity} if the following is true: given a sequence $\{\stmt_i\}_{i=1}^{\infty}$ such that $\stmt_i \narrower \stmt_j$ for all $ i \leq j$ and $\stmt = \bigOR_i\stmt_i$ the disjunction, for every $\stmt[t] \in \tdomain$ such that $\stmt_i\leq_{\stmt[u]} \stmt[t]$ for all $i$ we have $\stmt\leq_{\stmt[u]} \stmt[t]$.
@@
 
-We therefore want to make the following distinction. TODO: define countable vs uncountable
+We therefore want to make the following distinction.
 
 \begin{defn}
-	We say that $\stmt$ is a \textbf{uncountable} infinitesimal with respect to $\stmt[u]$ if all there exists a sequences $\{ \stmt_i \}_{i=1}^\infty$ such that:
+We say that $\stmt$ is a \textbf{countable infinitesimal} with respect to $\stmt[u]$ if it satisfies the following conditions:
 \begin{enumerate}
-		\item $\stmt \finer \stmt_i$ for all $i$
-		\item $\stmt_i \ncomp \stmt_j$ for all $i \neq j$
-		\item $\bigOR\limits_{i=1}^n \stmt_i \finer \stmt[u]$ 
+    \item $\stmt$ is an infinitesimal with respect to $\stmt[u]$
+    \item there exists a (countable) sequence $\{\stmt_i\}_{i=1}^{\infty}$ such that $\stmt_i\finer\stmt$ for all $i$, and $\bigOR\limits_{i=1}^{\infty}\stmt_i$ is comparable but not infinitesimal with respect to $\stmt[u]$
 \end{enumerate}
-	In words, $\stmt$ is a countable infinitesimal with respect to $\stmt[u]$ if we can find for any number, there is a disjunction of that many broader mutually incompatible statements $\stmt_i$ is finer than $\stmt[u]$.
+An infinitesimal which is not a countable infinitesimal will be called an \textbf{uncountable infinitesimal}. 
 \end{defn}
 
-TODO: desiderata that if there are no countable infinitesimal, then we have monotone continuity
-
-We wish to understand the following. This is monotone continuity up to infinitesimals, and so with the right definitions, it should translate to monotone continuity for $\lequ$.
-
-Let $\stmt[u]$ be a statement which does not admit countable infinitesimals, and let $\{\stmt_i:i\in\mathbb{N}\}$ be a sequence of statements monotone increasing in narrowness (i.e.
+An example of a space with countable infinitesimals is the natural numbers $\mathbb{N}$.
+If the possibilities are natural numbers themselves and they are equigranular, for any infinite subset $S$ of $\mathbb{N}$, each natural number is infinitesimal with respect to $S$.
+However, a countable sequence of natural numbers may not be infinitesimal with respect to $S$ (and indeed may surpass $S$ in coarseness).
+Another example of countable infinitesimals is any subset of $\mathbb{R}$ of finite positive Lebesgue measure.
+
+An example of a space with no countable infinitesimals with respect to the full space is an interval of finite length in $\mathbb{R}$ (any positive-Lebesgue measure subset will work).
+To see why, note that all infinitesimals are measure zero, and any countable union of measure zero sets is still measure zero.
+
+It seems possible that the only obstruction to monotone continuity is countable infinitesimals, since they allow limits of equigranular sequences to be of different granularity classes.
+
+\begin{desid}
+    If $\stmt[u]$ is such that there are no countable infinitesimals, then $\lequ$ has monotone continuity. 
+\end{desid}
+
+Before achieving the above, it may be possible to obtain a related result about fineness and monotone continuity, which could translate to monotone continuity for $\lequ$.
+
+\begin{desid}
+Let $\stmt[u]$ be a statement which does not admit countable infinitesimals, and let $\{\stmt_i\}_{i=1}^{\infty}$ be a sequence of statements monotone increasing in narrowness (i.e.
 $\stmt_i\narrower\stmt_j$ for all $i\leq j$).
@@
 If $\stmt[t]\in\tdomain$ such that $\stmt[t]$ is finitely comparable to $\stmt[u]$, and $\stmt_i\finer\stmt[t]$ for all $i$, then $\stmt\finer\stmt[t]\OR\stmt[w]$ for some infinitesimal (with respect to $\stmt[u]$) $\stmt[w]$.
+\end{desid}
+
+\section{Adapting results from \cite{villegas}}
+
+Tentative layout:
+- summarize the result in terms of input requirements (stated with our notation) and output
+- discussion of how far we are from getting those requirements
+- sketch of the main steps of the proof (the key propositions and the approach of the proof of each in our notation)
+- discussion on whether it is feasable and how to relax the proof for finite additivity
+
+The paper \cite{villegas} by Villegas gives what seems to be an extremely relevant overview of qualitative probability, including many of the axioms which we seek in our structures.
+
+In this part, we will assume that $\tdomain_{\stmt[u]}$ is a subset of the theoretical domain on which $\lequ$ is total, $\stmt[u]\in \tdomain_{\stmt[u]}$ is a distinguished element (we'll refer to it as the \emph{unit} when needed).
+We will use the word ``equiprobable" for $=_{\stmt[u]}$ since it means something different from equigranular (since $\leq_{\stmt[u]}$ is a different ordering from fineness, even though it will be derived from fineness).
+
+The goal is to use $\lequ$ as a qualitative probability.
+The first basic requirement is that $\leq_{\stmt[u]}$ is a total pre-order with maximal and minimal elements.
+These were contained in several desiderata earlier.
+
+Below is the version of the Axiom of Monotony for $\lequ$; earlier in Proposition \ref{fineness_monotony} it was shown that fineness satisfies this, and Definition \ref{smalleru} would allow this to descend to $\lequ$ as well.
+
+Axiom of Monotony: 
+Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
+Assume that $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, and $\stmt_2\ncomp \stmt_4$ (i.e.
+$\stmt_2\AND \stmt_4 =_{\stmt[u]} \impossibility$).
+Then we have $\stmt_1\OR \stmt_3 \leq_{\stmt[u]} \stmt_2\OR \stmt_4.$
+
+The following are consequences of $\lequ$ being a total pre-order with the Axiom of Monotony. Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
+\begin{itemize}
+    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ and $\stmt_2<_{\stmt[u]} \stmt_3$, then $\stmt_1<_{\stmt[u]}\stmt_3$. 
+    \item If $\stmt_1\narrower \stmt_2$ then $\stmt_1\leq_{\stmt[u]} \stmt_2$
+    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, $\stmt_2\leq_{\stmt[u]} \stmt_4$, and $v\leq_{\stmt[u]} \stmt_1$, then $\stmt_2 \AND \NOT \stmt_1 \leq_{\stmt[u]} \stmt_4 \AND \NOT \stmt_3$
+    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ then $\NOT \stmt_2 \leq_{\stmt[u]}\NOT \stmt_1$
+\end{itemize}
+
+Beyond the above, we will also utilize Monotone Continuity (Definition \ref{monotone_continuity}) in much of what follows.
+
+Another important notion for qualitative probability is the idea of atoms.
+
+\begin{defn}
+An \textbf{atom} is a statement $\stmt\in S$ such that $\impossibility<_{\stmt[u]}\stmt$ and there does not exist any statement $\stmt_2\in S$ such that $\impossibility<_{\stmt[u]} \stmt_2<_{\stmt[u]} \stmt$.
+\end{defn}
+In our framework, if possibilities are equigranular and not equivalent at $\stmt[u]$ scale to $\impossibility$, then they will be atoms.
+The presence or lack of atoms has several important consequences.
+In particular, Theorem 5 in \cite{villegas} \S3 includes the fact that $\tdomain_{\stmt[u]}$ is atomless if and only if every statement can be partitioned into two equiprobable statements.
+Theorem 3 in \S4 states that if $\tdomain_{\stmt[u]}$ is atomless, then there exists a unique probability measure compatible with $\leq_{\stmt[u]}$, and it is countably additive.
+
+A major goal of this work is to discover the right axioms and definitions that will lead to a notion of $\lequ$ which has the properties necessary for the existence (and uniqueness) of a compatible probability measure on $\tdomain_{\stmt[u]}$.
+The above axioms discussed are enough - several theorems from \cite{villegas} give the desired result (in particular, see Theorem 3 in Section 4 and Theorem 5(iv) in Section 3, as well as their proofs).
+However, we have not as of yet been able to prove that $\lequ$ has the right properties.
+It is also possible that our framework will allow us to make use of different sets of axioms that lead to a similar conclusion.
+The survey \cite{fishburnsurvey} contains some examples of different axioms which may be very helpful.
+
+\subsection{Old notes}
 
@@
 
-\subsection{Some thoughts on Example in 3.3}
-
-The construction above hints at some fundamental concepts in measure theory which will be important for us to consider.
-
-Notice that the two sets $V$ and $W$ are both of infinite Lebesgue measure (area).
-With respect to Lebesgue measure, monotone continuity is immediate since if $\bigcup\limits_0^n W\finer V$ for all $n$ then $\lim\limits_{n\to\infty}\mu(\bigcup\limits_0^n W) = \mu(W)\finer V$.
-Even the above example does not violate this, since there, both would be ``equivalent" in fineness, namely they are infinitely coarse.
-
-Our goal is to have a measure which is able to distinguish sets which may be of infinite Lebesgue measure as well.
-This will be very difficult to fully capture.
-Consider the following procedure.
-Starting with the vertical strip $V$, slide all the squares at even $y$-values to the right by one, and then allow the squares to collapse down to rest on the square below.
-Now, by simply rearranging regions, we have (almost) a cover of the wider strip $W$.
-The only missing points are the line $x=1, y>0$.
-If instead we used $(0,1]$ and $(0,2]$ for the strips, this procedure would cover them exactly.
-One could continue and cover the entire plane with similar translations.
-
-Here is one approach to distinguish $V$ and $W$, and similar methods will enable one to distinguish many more infinite-measure subsets.
-Let $S_n$ be the $2n\times 2n$ square centered the origin with vertices at $(\pm n, \pm n)$.
-Then for all $n$, $V\cap S_n$ has area $n$, and $W\cap S_n$ has area $2n$.
-This holds for all $n$ and so we might use something like this to think about how we would distinguish infinite areas.
-Notice that this is in some sense writing $V$ and $W$ as countable unions of nested sets in a standardized way (e.g.
-$V = \bigcup\limits_n V\cap S_n$) and then comparing the finite-area subsets.
-One could choose shapes other than $S_n$ for this purpose.
-Another way to distinguish them would be to look at the limit $\lim\limits_{n\to\infty}\text{Area}(V\cap S_n)/\text{Area}(S_n)$; in this case they are both 0 but for a different choice of $S_n$ one could find an asymptotic difference.
-The point of the above discussion is to try to gain a better idea of which (infinite/unbounded) sets we want to be related by $\finer$.
-It is not clear even for ``nice" regions of $\mathbb{R}^2$, let alone at the full generality we are seeking.
-
-Regardless, the goal is not just to distinguish infinite-Lebesgue-measure sets. We need to understand when monotone continuity is relevant and what happens without it.
+% END OF NOTES
 
@@
 We have $\bigcup\limits_0^\infty W_i = W$.
-This means $\bigcup\limits_{i=0}^n W_i\finer V$ for all finite $n$ but $\bigcup\limits_{i=0}^\infty W_i = W \coarser V$.
-
-TODO: show that this is a problem for $\lequ$
+This means $\bigcup\limits_{i=0}^n W_i\sfiner V$ for all finite $n$ but $\bigcup\limits_{i=0}^\infty W_i = W \scoarser V$.
+This is a violation of monotone continuity, including the definition given up to infinitesimals (since $\bigcup\limits_{i=0}^{\infty}W_i$ is more than an infinitesimal greater than $V$).
+
+By either definition of $\lequ$, this would give us an increasing sequence $W_n' := \bigcup\limits_{i=0}^n$, all of whose terms are (strictly) bounded by $V$, but whose limit (strictly) exceeds $V$ (in $\lequ$ terms), and so this violates monotone continuity of $\lequ$.
 
@@
 
-TODO: Make this an example of how measure zero is not equigranularity
+TODO(GC): Make this an example of how measure zero is not equigranularity
 
@@
 However, if we try to ``move" $\stmt$ to an equigranular copy inside $\stmt_1$, it will have infinitely many ``holes" - if we wish to use other points of $\stmt_1$ to fill them, there are numerous countable subsets which we consider non-equigranular, and so it is unclear if there is any canonical choice of equigranularity class for such an $\stmt_2$.
+
+Note that even two countable subsets of possibilities need not be equigranular.
+For example, if $\{\stmt[x]_i\}_{i=0}^{\infty}$ is a set of countably many (equigranular) possibilities, and $\stmt = \bigOR\limits_{i=0}^{\infty}\stmt[x]_i$, then $\stmt' = \bigOR\limits_{i=0}^{\infty}\stmt[x]_{2i}$ is strictly finer.
+
+Note: I didn't quite recall what we were thinking with this TODO.
+But if we are wanting to show that two measure zero sets can be non-equigranular, we can go much more generally and not rely on a measure with the above example of countable sets.
 
@@
 
+The problem is that infinite sets disrupt the relationship between narrowness and fineness, such that a maximal chain in narrowness is not necessarily a maximal chain in fineness.
+For example, if we have a countable set of possibilities (e.g.
+the natural numbers), a maximal chain in narrowness can start from the empty set, start by adding possibilities one be one until we end with the full set.
+The set with all possibilities except one, for example, cannot be added to the chain, but still represents a level of description that is in between all the finite sets and the full infinite set.
+The only case where we can construct maximal chains is precisely when the possibilities are finite, and is therefore not interesting.
+
 \begin{defn}
@@
 every pair of elements in $T$ is comparable in $\leq$).
-A chain is said to be \textbf{maximal} if it is not a proper subset of any other chain, i.e.
-there are no ``gaps" in the chain.
+A chain is said to be \textbf{saturated} if there are no ``gaps" in the chain.
+That is, there is no $s \in S$ such that $s \notin T$, $t_1 \leq s \leq t_2$ for some $t_1, t_2 \in T$ and $T \cup \{s\}$ is a chain.
 \end{defn}
-Observe that in a totally ordered set, there is only one maximal chain, namely the entire set.
+
+TODO(GC): everything needs to be changed to saturated
 
@@
 
-The issue is that the granularity of a countable sequence with each term a finite size but an infinite limit ``jumps'' at infinity.
-For example, take the sequence $\{1\}$, $\{1, 3\}$, $\{1, 3, 5\}$, ...
-, \{ all odd natural numbers\}.
-Compare to $\{1\}$, $\{1, 2\}$, $\{1, 2, 3\}$, ...
-, \{ all natural numbers\}.
-They are both maximal chains.
-They both go through all sizes of finite sets, but they reach infinity at two different levels.
-Effectively, the above property is only true over finite sets.
-
-The following consequences of the MCP are interesting, but only use much weaker statements about the fineness classes of possibilities.
+The following made the idea look promising.
 
@@
 
-TODO: add remark that MCP -> finite domain
-
-\begin{remark}
-The above observations and propositions show that MCP and comparability of possibilities imply equigranular possibilities, but equigrangular possibilities and totality do \emph{not} imply MCP.
-
-Further, MCP alone implies neither totality nor equigranular possibilities.
-Indeed, consider a theoretical domain with only two possibilities and a fineness ordering which has two unequal successors to the fineness of $\impossibility$.
-This satisfies the MCP but fails the other conditions.
-\end{remark}
+\begin{prop}
+Let $\tdomain$ be a theoretical domain with the maximal chains property. Then they must have finitely many possibilities.
+\end{prop}
+\begin{proof}
+Let $\tdomain$ be a theoretical domain with the maximal chains property and such that the set $X$ of possibilities is infinite.
+Let $X' = \{x_i\}_{i=0}^{\infty} \subseteq X$ be a countable subset.
+Let $s_j = \bigOR_{j=0}^{i} x_i$ be the disjunction of the first $j$ possibilities.
+Let $\stmt = \bigOR_{i=0}^{\infty} x_i$ be the disjuction of all the possibilities in the sequence.
+The set $C = \{ \impossibility, \stmt \} \cup \{\stmt_j\}_{j=0}^{\infty}$ is a chain.
+In fact $\impossibility \snarrower \stmt_j \snarrower \stmt$ for all $j$ and $\stmt_i \narrower \stmt_j$ if $i \leq j$.
+
+To show that $C$ is maximal, take a statement $\hat{\stmt} \narrower \stmt$ such that $\hat{\stmt} \notin C$.
+Let $A \subseteq X'$ be the set of possibilities compatible with $\hat{\stmt}$.
+If $A$ is infinite, since $\hat{\stmt} \nequiv \stmt$, there exists an $x_i \in X'$ such that $x_i \notin A$.
+Therefore $\stmt_{i} \nnarrower \hat{\stmt}$ since the first is compatible with a possibility that is not compatible with the second, and $\hat{\stmt} \nnarrower \stmt_{i}$ since $\hat{\stmt}$ is compatible with infinitely many statements and $\stmt_{i}$.
+If $A$ is finite, let the $x_i \in A$ for which $i$ is highest.
+Note that $A$ cannot contain all previous $x_j$ in the sequence, since $\hat{\stmt} \notin C$.
+Let $x_j$ the possibility with the highest $j \leq i$ that is not in $A$.
+We have $\hat{\stmt} \nnarrower \stmt_{j}$ and $\stmt_{i} \nnarrower \hat{\stmt}$ since each is compatible with a possibility incompatible with the other.
+Since we cannot add other statement, $C$ is maximal.
+
+Now we show that, even though $C$ is maximal, it does not go through all levels of granularity.
+Let $x \in X'$ and consider $\stmt \AND \NOT x$.
+This is finer than $\stmt$ and is coarser than all other statements statements in $C$ since they are all compatible with finitely many possibilities.
+Therefore MCP does not hold.
+Therefore, if MCP holds, the domain cannot have infinitely many possibilities.
+\end{proof}
 

### 2021-04-26 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
Update classical-phil-article-BJPS.tex
Cut most of section 6.2 (discussion of extended phase space formalism). Eliminated many footnotes and trimmed down other extraneous text (e.g. brief discussion of discrete case). Cut \geq 1185 words

@@
 A third condition is necessary for recovering Lagrangian mechanics.
-This shows that Lagrangian mechanics is a proper subset of Hamiltonian mechanics, elucidating the connection between these two formalisms.
+This shows that Lagrangian systems are a proper subset of Hamiltonian systems, elucidating the connection between these two formalisms.
 Using our framework, we argue that the invariant phase space density is physically fundamental.
 Since Hamiltonian mechanics provides the proper setting for this density, we gain another reason for privileging it over Lagrangian mechanics.
-Finally, we interpret Lagrangian mechanics within the Hamiltonian framework, providing a geometric interpretation of the principle of stationary action.
-We rebut arguments for privileging Lagrangian mechanics, showing where they go wrong.
+We interpret Lagrangian mechanics within the Hamiltonian framework, providing a geometric interpretation of the principle of stationary action.
+Finally, we rebut arguments for privileging Lagrangian mechanics, showing where they go wrong.
 \end{abstract}
@@
 In the context of comparing Hamiltonian and Lagrangian mechanics, three interpretive stances have emerged.
-\textcites[]{North} has argued for privileging Hamiltonian mechanics, on the grounds that it requires less mathematical structure to formulate.
+\textcites[]{North} has argued for privileging Hamiltonian mechanics on the grounds that it requires less mathematical structure to formulate.
 In contrast, \textcites[]{Curiel} recommends privileging Lagrangian mechanics, arguing that it isolates physically-significant kinematical constraints that all classical systems must satisfy.
 Finally, we can interpret Lagrangian and Hamiltonian mechanics as theoretically equivalent in a precise mathematical sense.
-Within a subclass of classical mechanical systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh}{Barrett2}.
+Within a subclass of systems known as the \textit{hyper-regular domain}, the two formalisms are categorically equivalent \parencites[]{Teh}{Barrett2}.
 Categorical equivalence provides a natural interpretive criterion for theoretical equivalence because it yields mutual inter-translatability between theory formulations.
@@
 We provide a geometric interpretation of the action principle, which applies even outside the hyper-regular domain.
-We also consider and rebut objections stemming from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
-Along the way, we examine the gauge redundancy of both the conjugate momentum (Section~\ref{gauge_momentum}) and the Lagrangian (Section~\ref{gauge}).
+We end by rebutting objections that arise from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
+%Along the way, we examine the gauge redundancy of both the conjugate momentum (Section~\ref{gauge_momentum}) and the Lagrangian (Section~\ref{gauge}).
 
@@
 In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by `fundamental.' Of the various notions of fundamentality, we will mean something entirely non-mysterious: a formalism is more fundamental provided that it is logically more general.
-In this case, the Lagrangian framework is less general because it requires that an additional condition hold, compared to the Hamiltonian framework.
+In this case, the Lagrangian framework is less general because it requires an additional condition to hold, compared to the Hamiltonian framework.
 Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.\footnote{One might worry that our criterion for fundamentality is subject to elementary counterexamples.
@@
 
-We cannot always determine the evolution of a system in phase space from its trajectory in physical space, since some systems fail to satisfy the condition of motion equivalence.
+We cannot always determine the evolution of a system in phase space from its trajectory in physical space, since some systems fail to satisfy motion equivalence.
 In general, knowing a system's position and velocity does not suffice for knowing its momentum.
@@
 
-Of course, different definitions of `Hamiltonian' and `Lagrangian systems' can result in different interpretive relationships.
+Of course, different definitions of `Hamiltonian' and `Lagrangian' systems can result in different interpretive relationships.
 On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems satisfy the motion equivalence assumption.
@@
 Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
-For details of the Lagrangian approach to dissipative systems, see \textcites[]{Smith}.} Hence, our claim about the fundamentality of Hamiltonian mechanics is restricted to the non-dissipative regime.
-We will view non-dissipative systems as neither Lagrangian nor Hamiltonian.
-Finally, our treatment is not restricted to $n$-particle mechanics but applies as well to rigid-body and continuum mechanics.\footnote{For discussion of the numerous conceptual problems that arise within different approaches to classical mechanics (including point-masses, rigid-bodies, and continua) see \textcites{Wilson}.
-In Wilson's terminology, the framework we apply here may be yet another `theory facade' for classical physics, but a fruitful one nonetheless.} This supports our contention that the property of infinitesimal reducibility fruitfully characterizes classical systems.
-Before presenting the derivational details of our argument in Section~\ref{derivation}, we clarify what we mean by `fundamental.'
+For details of the Lagrangian approach to dissipative systems, see \textcites[]{Smith}.} This restricts our claim about the fundamentality of Hamiltonian mechanics to the non-dissipative regime.
+Finally, our treatment is not restricted to $n$-particle mechanics but applies as well to rigid-body and continuum mechanics.
+This supports our claim that the property of infinitesimal reducibility fruitfully characterizes classical systems.
+Before presenting the derivational details of our argument in Section~\ref{derivation}, we describe how we avoid central problems facing North's privileging of Hamiltonian mechanics.
+%We view non-dissipative systems as neither Lagrangian nor Hamiltonian. 
+%\footnote{For discussion of the numerous conceptual problems that arise within different approaches to classical mechanics (including point-masses, rigid-bodies, and continua) see \textcites{Wilson}.
+In Wilson's terminology, the framework we apply here may be yet another `theory facade' for classical physics, but a fruitful one nonetheless.}
 
@@
 Compared with North's \parencites*[]{North} account of structure, our definition has two advantages.
-First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which empiricists and pragmatists decry.
+First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which scientific anti-realists and pragmatists decry.
 Secondly, it avoids a technical objection that Swanson and Halvorson \parencites*[]{Swanson} levy against North's more precise mathematical criterion for comparing structure.
@@
 
-Owing to our epistemic inabilities to access these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant densities (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus.
+Owing to the epistemic difficulty of accessing these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant densities (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus.
 Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration.
@@
 Regardless of its metaphysical status, Hamiltonian mechanics is physically and conceptually more fundamental.
-Indeed, our argument does not privilege any one of the various mathematical formalizations of Hamiltonian mechanics.\footnote{These include formulating Hamiltonian mechanics on a $2n+1$-dimensional contact manifold, on an extended phase space, on a Poisson manifold, and the standard formulation on a symplectic manifold that we typically use below.} Each of these frameworks has different pragmatic costs and benefits.
-Hence, it is plausible that there may be no single, uniquely necessary mathematical structure for mechanics, but rather a family of related structures.
-In other words, there may be no canonical minimal structure for formalizing the relevant physical models and physical assumptions.\footnote{Tulczyjew's reformulation of Hamiltonian mechanics further supports viewing symplectic geometry---and hence Hamiltonian mechanics---as conceptually fundamental.
-According to Teh and Tsementzis, this reformulation shows that  ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry'' \parencites*[46]{Teh}.}  Additionally, each of these formalisms references some `unphysical' coordinates or representational means, whose physical significance is borne out by their fruitfulness in modeling.
+Indeed, our argument does not privilege any one of the various mathematical formalizations of Hamiltonian mechanics.
+These include formulating Hamiltonian mechanics on a $2n+1$-dimensional contact manifold, on an extended phase space, on a Poisson manifold, and the standard formulation on a symplectic manifold that we recover in Section~\ref{deterministic}.\footnote{Tulczyjew's reformulation of Hamiltonian mechanics further supports viewing symplectic geometry---and hence Hamiltonian mechanics---as conceptually fundamental.
+According to Teh and Tsementzis, this reformulation shows that  ``the very concept of a classical mechanical system (and not merely the phase space of that system) should be described in terms of symplectic geometry'' \parencites*[46]{Teh}.}
 
@@
 
-Advantageously, our criterion for comparing Hamiltonian and Lagrangian mechanics avoids the criticisms that Swanson and Halvorson develop against North.
+Advantageously, our criterion for comparing Hamiltonian and Lagrangian mechanics avoids these problems.
 Hamiltonian mechanics has less structure than Lagrangian mechanics simply in the sense of logical strength: a system must satisfy strictly fewer physical conditions in order to be Hamiltonian.
-Whereas to be Lagrangian, the system must also satisfy the assumption of motion equivalence.
-The assumption of motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.
+Whereas to be Lagrangian, the system must also satisfy the motion equivalence condition.
+Motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.
 Section~\ref{density} provides a further argument for viewing Hamiltonian mechanics as more fundamental.
@@
 \textit{Infinitesimal reducibility}: the state of the system is reducible to the state of its infinitesimal parts.
-That is, giving the state of the whole system is equivalent to giving the state of its parts, which is equivalent to giving the states of those parts, and so on.
+That is, specifying the state of the whole system is equivalent to specifying the state of its parts.
 \es
@@
 
-If $\mathcal{S}$ is a discrete space with $U \subseteq \mathcal{S}$, we define $\mu(U) = \sum\limits_{s \in U} \rho_c(s)$ so that $\mu(\mathcal{S})$ gives us the size of the system and $\mu(U)/\mu(\mathcal{S})$ gives us the fraction of the particles that can be found in the region $U$.\footnote{The discrete case is far less interesting because discrete quantities cannot change continuously over time, meaning that they must be constant.
+%If $\mathcal{S}$ is a discrete space with $U \subseteq \mathcal{S}$, we define $\mu(U) = \sum\limits_{s \in U} \rho_c(s)$ so that $\mu(\mathcal{S})$ gives us the size of the system and $\mu(U)/\mu(\mathcal{S})$ gives us the fraction of the particles that can be found in the region $U$.\footnote{The discrete case is far less interesting because discrete quantities cannot change continuously over time, meaning that they must be constant.
 We discuss this elsewhere.}
@@
 
-In the continuous case, $\mathcal{S}$ is a manifold, and we define $\mu(U) = \int_U \rho_c(s) d\mathcal{S} $.
-We will call a set of quantities $\xi^a$ \textit{state variables} provided that they fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates' and $\xi$ is a coordinate map.
+When physical quantities change continuously over time, $\mathcal{S}$ is a manifold, and we define $\mu(U) = \int_U \rho_c(s) d\mathcal{S} $.
+State variables $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates' and $\xi$ is a coordinate map.
 We will have multiple manifolds, so to avoid confusion, we will call the coordinates of the state space `state variables'.
 As usual with coordinate charts, $\xi$ is properly defined on some $U \subseteq \mathcal{S}$, and $\xi^{-1}$ is defined on $\xi(U) \subseteq \mathbb{R}^n$.
-We will neglect these subtleties in what follows.} In other words, $\xi$ represents a state $s \in \mathcal{S}$ by a set of state variables $\xi^a \in \mathbb{R}^n$, such that $\xi (s) = \xi^a$.
+We will neglect these subtleties in what follows.} The map $\xi$ represents a state $s \in \mathcal{S}$ by a set of state variables $\xi^a \in \mathbb{R}^n$, such that $\xi (s) = \xi^a$.
 To represent the distribution $\rho_c$ in terms of state variables, we define a density $\rho_{c, \xi}$: $\mathbb{R}^n \to \mathbb{R}$ by $\rho_{c, \xi} \equiv (\rho_c \circ \xi^{-1})$.
@@
 Likewise, in the numerator of the Jacobian, we denote $\partial \phi$ by $\partial \hat{\xi}^b$, where $\hat{\xi}^b$ is here understood as a function of $\xi^a$.} Here, $ \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$ is the determinant of the Jacobian for this change of state variables.
-From the preceding two equations, we see that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.
-Matching integrands, we have $\rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$.
+Matching integrands from the preceding two equations, we have $\rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|$.
 From the scalar transformation property, $\rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
@@
 We will now show that in order for the density $\rho_{c, \xi}$ to have these transformation properties, the classical phase space must be at least a symplectic manifold.
+%From the preceding two equations, we see that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.
 
@@
 Two unit variables are \textit{independent} if one unit variable can be changed without changing the other.
-For example, one can measure the vertical position in meters while measuring the horizontal position in miles.
-In general, the unit variables $q^i$ are the subset of the state variables that define the unit system, which is needed to relate the properties of the system to the measured values.
+For example, one can measure vertical position in meters while measuring horizontal position in miles.
+In general, the unit variables $q^i$ are the subset of the state variables that define the unit system, which relates the properties of the system to measured values.
 The $q^i$ determine the units not only for themselves, but for all other state variables.
-A \textit{change of units} is therefore a map $\hat{q}^j = \hat{q}^j(q^i)$ that defines new unit variables in terms of the old unit variables alone.
+A \textit{change of units} is therefore a map $\hat{q}^j = \hat{q}^j(q^i)$ that defines new unit variables in terms of old unit variables.
 Since the unit variables define the units for all state variables, this change must induce a unique change of state variables $\hat{\xi}^b = \hat{\xi}^b(\xi^a)$ for all variables.
@@
 Let us consider the simplest case: suppose that a single unit variable $q $ suffices to define the unit system for the particle state space.
-Then, the state variables are given by a set $\xi^a = \{ q, k^\alpha \}$, where \textit{prima facie} the index $\alpha$ could be any non-negative integer.
+Then, the state variables are a set $\xi^a = \{ q, k^\alpha \}$, where \textit{prima facie} the index $\alpha$ could be any non-negative integer.
 We will show that, in fact, there must be exactly one $k^\alpha$ variable, conjugate to the single $q$ variable.
 
-As a scalar, the units of the density $\rho_{c, \xi}$ are invariant under changing the units of the state variables.
-So suppose we change the units by transforming $q$ to a new unit variable $\hat{q}=\hat{q}(q)$.
+As a scalar, the units of the density $\rho_{c, \xi}$ are invariant under unit changes of the state variables.
+Suppose we change units by transforming $q$ to a new unit variable $\hat{q}=\hat{q}(q)$.
 Call the new state variables $\hat{\xi}^b = \{ \hat{q}, \hat{k}^\beta\}$.
@@
 Hence, if there were three or more $\xi^a$ variables, this constraint would be insufficient to recover the transformation of the $k^\alpha$ uniquely.
-Yet in this case, $\hat q$  would not fully define the units for all other state variables, contrary to assumption.
+In this case, $\hat q$  would not fully define the units for all other state variables, contrary to assumption.
 Consequently, there must be exactly two state variables: $q$ and a single $k$.
@@
 Likewise, transforming the second unit-variable $q^2$ in the same way (while fixing $k_1$) requires a corresponding $k_2$.
-In the end, we will find that $\mathcal{S}$ is $2n$-dimensional, and the state variables are $\xi^a = \{ q^i, k_i \}$.
+Continuing, we find that $\mathcal{S}$ is $2n$-dimensional, with state variables $\xi^a = \{ q^i, k_i \}$.
 We define an \textit{independent degree of freedom} as the space charted by a pair of such variables.
@@
 Consider a parallelepiped, whose opposite sides have equal areas.
-Taking the integral over this closed surface (over the parallelepiped), opposite sides provide equal and opposite contributions.
+Integrating over the closed surface of the parallelepiped, opposite sides provide equal and opposite contributions.
 Hence, the integral over this closed volume must equal zero.
@@
 Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
-Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$---at the initial time $t_0$---uniquely identifies the state of the system.
+Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$ at time $t_0$ uniquely identifies the state of the system.
 Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial state at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e.
-the density remains the same throughout the evolution.\footnote{For convenience, we will sometimes drop the composite state $c$ and coordinate-choice $\xi$ indices from the invariant density $\rho_{c, \xi} (\xi^a)$.
-We will reserve the word `distribution' for the coordinate-free distribution $\rho_c (s)$.
+the density remains the same throughout the evolution.\footnote{For convenience, we sometimes drop the composite state $c$ and coordinate-choice $\xi$ indices from the invariant density $\rho_{c, \xi} (\xi^a)$.
+We reserve the word `distribution' for the coordinate-free distribution $\rho_c (s)$.
 } In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa.
@@
 Now consider the integral $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
-The region $\Sigma$ will be mapped to a new region, but the fraction of the system found in both the old and new regions must be the same.
+Mapping the region $\Sigma$ to a new region, the fraction of the system found in both the old and new regions must be the same.
 Since both the integral and $\rho_c$ must remain constant during the evolution, the two-form $\omega$ must also remain the same.
-That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (otherwise, volumes would not be mapped to equal volumes, which would violate determinism).
-The evolution is therefore a symplectomorphism and corresponds to Hamiltonian evolution.
+That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (otherwise, volumes would not be mapped to equal volumes, violating determinism).
+The evolution is therefore a symplectomorphism, corresponding to Hamiltonian evolution.
 
@@
 Since these trajectories are differentiable in $x^i$, we can define a velocity $v^i = d_t x^i$.
-If the equations of motion were such that the velocity were a function of only the unit-variables (i.e.
+If the equations of motion contained the velocity as a function of only the unit-variables (i.e.
 $v^i=v^i(q^i)$), then motion equivalence would fail.
-For if both $x^i$ and $v^i$ were a function of only $q^i$, this relationship would not be invertible because of dimensional considerations (a surjective function from an $n$-dimensional manifold to a $2n$-dimensional manifold cannot be invertible).
+For if both $x^i$ and $v^i$ were functions of only $q^i$, this relationship would not be invertible due to dimensional considerations (a surjection from an $n$-dimensional manifold to a $2n$-dimensional manifold cannot be invertible).
 Hence, we must have $v^i=v^i(q^j, p_k)$, so that knowledge of the state in phase space suffices for knowledge of the trajectory.
 If motion equivalence holds, then this relationship must be invertible, meaning that knowledge of the trajectory suffices for knowledge of the phase space state.
-For systems that satisfy the motion equivalence condition, then, the following relationships hold at any given time:
+For systems that satisfy the motion equivalence condition, the following relationships thus hold at any given time:
 \begin{equation}\label{weak_equivalence}
@@
 In this section, we argue that the special relationship between Hamiltonian mechanics and invariant densities provides a further reason to privilege Hamiltonian mechanics.
-Additionally, we argue that viewing physics through the lens of invariant densities has much to commend it.
-In general, it provides a better understanding of classical systems.\footnote{It also provides a better understanding of statistical mechanics and entropy, since phase space provides a uniform measure over q and p.
+Additionally, we argue that viewing physics through the lens of invariant densities provides a better understanding of classical systems.\footnote{It also provides a better understanding of statistical mechanics and entropy, since phase space provides a uniform measure over q and p.
 We discuss this elsewhere.}
@@
 The phase space density $\rho$ is physically fundamental because it tracks how much of the system is in each possible state.
-It quantifies the amount of stuff in each phase space state, in virtue of representing the distribution.
-Typically, the canonical position and momentum observables play this role, since they suffice to completely specify a classical state; they fix the values of all other observables.
+By representing the distribution, the density quantifies the amount of stuff in each phase space state.
+Typically, the canonical position and momentum observables play the role of specifying a classical state, since they fix the values of all other observables.
 Ruetsche refers to this as ``fundamental in the physicist's sense'' \parencite*[31, 200]{Ruetsche}.
-However, we will show in Section~\ref{gauge_momentum} that the conjugate momentum is gauge-redundant.
-For this reason, the invariant density provides a more fundamental specification of the state than the state variables themselves.\footnote{Section~\ref{infinitesimal} provides another reason for viewing the density as fundamental: by defining the density $\rho_{c, \xi}$, we recover the right mathematical structure, namely a symplectic manifold.
+However, we will show in Section~\ref{gauge_momentum} that the conjugate momentum is gauge-dependent.
+For this reason, the invariant density provides a more fundamental specification of the state than the state variables themselves.
+%\footnote{Section~\ref{infinitesimal} provides another reason for viewing the density as fundamental: by defining the density $\rho_{c, \xi}$, we recover the right mathematical structure, namely a symplectic manifold.
 If we did not define the density, then we would need another account of why this mathematical structure arises for classical systems.}
@@
 This gives a further reason to privilege Hamiltonian mechanics.
-Section~\ref{gauge_momentum} shows why the conjugate momentum $p$ is itself a gauge-dependent quantity.
+Section~\ref{gauge_momentum} shows why the conjugate momentum $p$ is gauge-dependent.
 This bolsters our claim that the density is more fundamental than the choice of state variable pairs.
@@
 If $\rho(x, y)$ is a function of multiple variables, dimensional analysis requires that the units of $\rho$ are consistent with the units of the arguments in the expression $\int_U \rho(x, y) dx dy$.
-The issue is that a manifold, in general, does not provide us with a way to measure areas and volumes.
+Yet in general, a manifold does not provide a way to measure areas and volumes.
 For example, a manifold may capture the physical characteristics of a group of people in terms of their height $h$ and their weight $w$, leading to a density $\rho(h,w)$ with units of people/(m$\cdot$Kg).
@@
 In the above example, $\rho(h,w) = |J| \rho(h, bmi) = \frac{1}{h^2} \rho(h, bmi)$.
-For a more familiar example, changing from polar to Cartesian coordinates, we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
+Similarly, in changing from polar to Cartesian coordinates we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
 Since the Jacobian is a function of the point, the value of $|J|$ can change from point to point.
@@
 The result is that we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in (in the sense of space-time coordinates).
-Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space.
-See Section~\ref{gauge} for discussion.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.\footnote{One might wonder if it is possible to construct a coordinate-independent density, perhaps using either i) volume forms from differential geometry or ii) the canonical volume form of Riemannian geometry.
+Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space $(q^i, p_i, t, E)$.
+See \textcites[]{Struckmeier}.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.\footnote{One might wonder if it is possible to construct a coordinate-independent density, perhaps using either (i) volume forms from differential geometry or (ii) the canonical volume form of Riemannian geometry.
 Both of these strategies fail, although space precludes a demonstration.}
@@
 
-Before proceeding, we provide a precise account of what gauge-dependency physically represents.
-In the spirit of our previous constructions, we consider a physically-motivated definition that leads to the correct mathematical structure.
-First, note that when motion equivalence holds, we can completely characterize states in terms of the position $q$ and its first derivative $\dot{q}$.
+We begin with a precise account of what gauge-dependency physically represents.
+As before, we consider a physically-motivated definition that leads to the correct mathematical structure.
+When motion equivalence holds, we can completely characterize states in terms of the position $q$ and its first derivative $\dot{q}$.
 Hence, any canonical transformation in phase space that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant.
 We will show that these transformations are exactly the gauge transformations.
-
-Since the position does not change, the new state variables can be written as $(q, \hat{p})$. Since the transformation is canonical, the new state variables must respect the usual Poisson brackets:
+Since the position does not change, the new state variables can be written as $(q, \hat{p})$.
+They must respect the usual Poisson brackets because the transformation is canonical:
 \begin{equation}
@@
 \end{equation}
-where $A_j$ are arbitrary functions of position. Using the third Poisson bracket, we have:
+where the $A_j$ are arbitrary functions of position.
+
+Using the third Poisson bracket, we have:
 \begin{equation}
@@
 Insofar as coordinate-invariance is a plausible necessary condition for fundamental physical status, so is gauge-invariance.
-By being gauge-dependent, the conjugate momentum cannot be physically fundamental, although it remains physically significant (in the same way that coordinate-variant quantities can be physically significant but not fundamental).
+By being gauge-dependent, the conjugate momentum cannot be physically fundamental.
+Nevertheless, it remains physically significant, similar to how coordinate-variant quantities can be physically significant but not fundamental.
 %might not need this paragraph; tacitly references the section 5.2 I deleted.
@@
 By analyzing classical systems in terms of their invariant phase space density, we gain new insights into Lagrangian mechanics.
-Section~\ref{action} shows how the action principle can be understood geometrically in the extended phase space as a consequence of the conserved density.
-We show that the action principle holds even when the action cannot be expressed as a function of just position and velocity, since it can always be expressed as a function of position and conjugate momentum.
-Section~\ref{gauge} discusses how the Lagrangian is gauge-dependent, and therefore less physical than the Hamiltonian.
-This provides a further argument for privileging Hamiltonian mechanics in the interpretation of classical systems.
-Finally, Section~\ref{Curiel} considers Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics, which he alleges encodes the necessary and sufficient structure for capturing the physically significant kinematical constraints of classical systems.
+Section~\ref{action} interprets the action principle geometrically in extended phase space as a consequence of the conserved density.
+We show that the action principle holds even when the action is not defined as a function of position and velocity, since it can always be expressed as a function of position and conjugate momentum.
+Our construction makes the gauge-dependence of the Lagrangian manifest.
+Section~\ref{Curiel} considers Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics, which he alleges encodes the necessary and sufficient structure for capturing the physically significant kinematical constraints of classical systems.
 This raises a number of objections to our own position, which we rebut.
@@
 Although central to Lagrangian mechanics, we can actually derive the stationary action principle using Hamiltonian mechanics.
-Our derivation simply assumes that the system trajectory follows a deterministic and reversible evolution.
+Our derivation simply assumes that the system's trajectory follows a deterministic and reversible evolution.
 We present the derivation for the one-dimensional case---where it is most easily visualized---but it can be generalized using differential geometry.\footnote{For a related argument, see \textcites[236--37, 243--45]{Arnold}.}
@@
 Again using the gauge freedom of $\vector{S}$, we can set $g(q)=0$, so that $\theta_q = p$.
-Thus, so far, we have shown that $\vector{\theta} = (p, 0,\theta_t) $.
+Thus far, we have shown that $\vector{\theta} = (p, 0,\theta_t) $.
 
@@
 If motion equivalence does not hold, we can define the action more generally over the function $L_p \equiv p \dot{q} - H(q, p, t)$, considering paths in the extended phase space.
-By deriving the stationary action principle in this more general setting, we simultaneously derive its usual version within Lagrangian mechanics.\footnote{When motion equivalence holds, the Legendre transform is well-defined, allowing us to express $L_p$ as a function of $(q, \dot{q}, t)$, i.e.
-$L_p = L(q, \dot{q}, t) = p \dot{q} - H$.} We have then that $ \mathscr{A}[\gamma] = \integral^{t_1}_{t_0} L_p (q, p, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$, where $\gamma$ is an actual trajectory followed by the system.
+By deriving the stationary action principle in this more general setting, we simultaneously derive its usual version within Lagrangian mechanics.
+We have then that $ \mathscr{A}[\gamma] = \integral^{t_1}_{t_0} L_p (q, p, t) dt = \integral^{t_1}_{t_0} [p \frac{dq}{dt} - H(q, p, t) ]dt$, where $\gamma$ is an actual trajectory followed by the system.
 Multiplying the Hamiltonian by $dt/dt = 1$ and adding a vanishing $0 \cdot \frac{dp}{dt}$, we obtain $ \mathscr{A}[\gamma] = \integral^{t_1}_{t_0} [p \frac{dq}{dt} + 0 \cdot \frac{dp}{dt} - H(q, p, t) \frac{dt}{dt} ]dt$.
@@
 To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
+%\footnote{When motion equivalence holds, the Legendre transform is well-defined, allowing us to express $L_p$ as a function of $(q, \dot{q}, t)$, i.e. $L_p = L(q, \dot{q}, t) = p \dot{q} - H$.}
 
@@
 This establishes a more general Hamilton's principle of stationary action, which we have derived simply by assuming that the displacement vector is divergenceless.
-In the special case that motion equivalence holds, the Legendre transformation is well-defined and $L_p = p \dot{q} - H(q, p, t)$ equals the usual Lagrangian $L(q, \dot{q}, t)$, thereby recovering the usual principle of stationary action for Lagrangian mechanics.
-
-\subsection{Gauge redundancy of the Lagrangian}
-\label{gauge}
-
-As discussed in Section~\ref{gauge_momentum}, the conjugate momentum $p$ is a gauge-dependent variable.
-In contrast, position and velocity are both gauge-invariant, making them physically privileged.
-Consequently, Lagrangian mechanics employs state space variables that are gauge-invariant, while Hamiltonian mechanics uses a gauge-dependent variable.
-Despite this advantage, Lagrangian mechanics has a compensating disadvantage: the Lagrangian is itself gauge-dependent.
-The formalism of Section~\ref{action} makes this gauge-dependence manifest.
-Representing the Lagrangian as $L =\vector{\theta} \cdot \vector{S} $, we see that since the vector potential $\vector{\theta}$ is gauge-dependent while the displacement vector $\vector{S}$ is not, $L$ must also be gauge-dependent.\footnote{As shown in Section~\ref{action} we can arbitrarily transform $\vector{\theta}$ by adding the gradient of an arbitrary scalar function, while $\vector{S}=curl(-\vector{\theta}) $ remains invariant under this gauge transformation.} The gauge-dependence of the Lagrangian makes it difficult to provide a clear physical interpretation of any particular Lagrangian.\footnote{The action inherits the gauge-dependence of the Lagrangian.
-This does not affect the principle of stationary action because the local minima and maxima of the action remain invariant under gauge transformations.} In contrast, the Hamiltonian is defined up to a constant, endowing it with greater physical significance.
-Overall then, we see that neither Lagrangian nor Hamiltonian mechanics has the upper hand when it comes to invoking only gauge-independent physical quantities.
-Both formalisms have their trade-offs in this regard (although, it is worth emphasizing that---as argued in Section~\ref{density}---Hamiltonian mechanics provides the natural setting for the invariant phase space density).
-
-Moving from gauge transformations to coordinate transformations, we gain further reasons for using Hamiltonian mechanics within the extended phase space formalism.
-For many applications, it suffices to use time-independent coordinate transformations.
-In this special case, both the Lagrangian $L$ and the Hamiltonian $H$ are coordinate-invariant.
-However, some applications benefit from (or even require) the use of time-dependent coordinate transformations, e.g.
-a rotating frame of reference.
-In this case, the Lagrangian remains coordinate-invariant, while the Hamiltonian becomes coordinate-dependent.
-Again, the formalism of Section~\ref{action} makes this time-dependence manifest.
-As we saw there, the Hamiltonian $H$ is the time component of the vector potential $\vector{\theta} $, entailing that $H$ transforms as the time component of a co-vector.
-Methodologically, this makes the Hamiltonian framework less convenient for treating this class of problems.
-
-Nonetheless, reformulating Hamiltonian mechanics within the extended phase space $(q^i, p_i, t, E)$ provides one way to rectify these deficiencies.\footnote{For a detailed discussion of the extended phase space formalism, see \textcites[]{Struckmeier}.} At root, the time-dependence of the Hamiltonian stems from the dual role of time as both a coordinate and a time-like parameter within the usual Hamiltonian framework.
-Yet, differential geometry recommends that we separate these two roles.
-We thereby add a temporal degree of freedom $(t, E) $, expanding the phase space degrees of freedom by one (recall that we previously had one degree of freedom for every canonical coordinate pair $q^i, p_i$).
-Next, we define the phase space evolution over an evolution parameter $s $.
-This yields two more equations of motion: $dt/ds = -\partial \mathcal{H}/ \partial E$ and $dE/ds = -\partial \mathcal{H}/ \partial t$, where $\mathcal{H}(q^i, p_i, t, E)$ is the extended Hamiltonian.
-The extended Hamiltonian not only acts as a generator of the evolution parameter $s$, but it also provides us with the \textit{Hamiltonian constraint} $\mathcal{H}=0$.
-The Hamiltonian constraint provides the relationship between the energy and the other state variables.
-Crucially, the extended Hamiltonian---unlike the ordinary Hamiltonian---is coordinate-invariant, even with respect to time-dependent coordinate transformations.
-For example, if we define the evolution parameter as the proper time, then the extended Hamiltonian is proportional to the mass (which is manifestly coordinate-invariant).
-This provides a clear physical meaning for the extended Hamiltonian.
-We have now considered many conceptual and physical advantages of the Hamiltonian formalism.
-In the next section, we rebut Curiel's \parencites*[]{Curiel} arguments against privileging Hamiltonian mechanics.
+In the special case that motion equivalence holds, the Legendre transformation is well-defined and $L_p = p \dot{q} - H(q, p, t)$ equals the usual Lagrangian $L(q, \dot{q}, t)$, thereby recovering the principle of stationary action for Lagrangian mechanics.
+
+As a corollary, this construction shows that the Lagrangian $L =\vector{\theta} \cdot \vector{S} $ is manifestly gauge-dependent.
+The vector potential $\vector{\theta}$ is gauge-dependent because we can transform it by adding the gradient of an arbitrary scalar function.
+The displacement vector $\vector{S} = curl(-\vector{\theta})$ is invariant under this gauge transformation.
+Hence, $L$ must be gauge-dependent, which complicates the physical interpretation of any particular Lagrangian.
+In contrast, the Hamiltonian is defined up to a constant, endowing it with greater physical significance.
 
@@
 Curiel contends that this form is the ``necessary and sufficient condition [a Hamiltonian must have] for it to model an abstract classical system'' \parencites*[305]{Curiel}.
-Since many possible Hamiltonians do not take this form,\footnote{As an example, Curiel considers the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$  \parencites*[305]{Curiel}.} Curiel believes we must artificially restrict Hamiltonian mechanics to those that satisfy Equation~\ref{6.2}.
+Since many possible Hamiltonians do not take this form,\footnote{As an example, Curiel considers the Hamiltonian $H = \frac{1}{2} p_i p^i + \sum_j p_j$  \parencites*[305]{Curiel}.} Curiel believes we must artificially restrict Hamiltonian mechanics to those systems that satisfy Equation~\ref{6.2}.
 
@@
 Given an initial probability distribution over states, the Lagrangian fails to specify how this distribution will evolve in time.
-In contrast, a Hamiltonian still defines a unique solution (i.e.
-a unique evolution in phase space) even when outside the hyper-regular domain (i.e.
-when the Hessian is zero).\footnote{Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
-There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other'' \parencites*[1179]{Barrett2}.} For instance, a zero Hamiltonian describes a static system (by Hamilton's equations, the derivatives of the position and momenta are zero).
+In contrast, a Hamiltonian still defines a unique solution---a unique evolution in phase space---even when outside the hyper-regular domain, i.e.
+when the Hessian is zero.
+For instance, a zero Hamiltonian describes a static system (by Hamilton's equations, the derivatives of the position and momenta are zero).
 Indeed, Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside the hyper-regular domain \parencites*[291, 308]{Curiel}.
-Curiel argues this favors Lagrangian mechanics, but it generates challenges for understanding non-hyper-regular Lagrangians.\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.
+Curiel argues this favors Lagrangian mechanics, but it generates challenges for understanding non-hyper-regular Lagrangians.
+%\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.
 In the case where a system is non-regular only at certain points, we can decompose the system into hyper-regular subsystems, each with its own Lagrangian (or Hamiltonian).}
+%\footnote{Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
+There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other'' \parencites*[1179]{Barrett2}.}
 
@@
 Additionally, our definition accommodates models of classical massless particles by maintaining the uniqueness of solutions, even outside the hyper-regular domain.
-Moving beyond classical mechanics, our framework has the advantage of providing an elegant relationship with quantum mechanical systems.
+Moving beyond classical mechanics, our framework has the advantage of providing an elegant relationship with quantum mechanics.
 By changing the requirement of infinitesimal reducibility to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.\footnote{We provide details of the quantum case elsewhere.} Hence, our framework perspicuously distinguishes quantum from classical systems.
@@
 By focusing on three basic physical conditions, we have derived the mathematical structures that underpin Hamiltonian and Lagrangian mechanics.
-These three physical conditions are i) infinitesimal reducibility, (ii) deterministic and reversible evolution, and (iii) motion equivalence.
+These three physical conditions are (i) infinitesimal reducibility, (ii) deterministic and reversible evolution, and (iii) motion equivalence.
 Each condition gives rise to corresponding mathematical structure, allowing us to interpret this structure in a physically well-motivated fashion.
@@
 Likewise, our construction provides a principled response to Curiel's \parencites*[]{Curiel} criticism that Hamiltonian mechanics requires an \textit{ad hoc} condition.
-This argumentative core already provides compelling reasons for interpreting Hamiltonian mechanics as more fundamental than Lagrangian mechanics, despite the fact that both frameworks adequately treat hyper-regular systems.
-
-Besides this, our approach has many additional interpretive payoffs.
+Although both frameworks adequately treat hyper-regular systems, there are compelling reasons to interpret Hamiltonian mechanics as more fundamental than Lagrangian mechanics.
 Our formalism emphasizes the fundamental significance of the invariant phase space density, along with the role of the symplectic form in counting states for each degree of freedom.
 Whereas the state variables $(q, p)$ suffer from the gauge-dependency of the conjugate momentum, the density provides an invariant description within the Hamiltonian formalism.
-Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain.
-Our construction also makes the gauge-dependency of the Lagrangian manifest.
+Finally, we have provided a geometrical interpretation of the action principle that makes the gauge-dependency of the Lagrangian manifest.
 Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but classical systems in general.

### 2021-04-27 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
ClassMechPhil: minor typos

@@
 This shows that Lagrangian systems are a proper subset of Hamiltonian systems, elucidating the connection between these two formalisms.
-Using our framework, we argue that the invariant phase space density is physically fundamental.
+Using our framework, we furthermore argue that the invariant phase space density is physically fundamental.
 Since Hamiltonian mechanics provides the proper setting for this density, we gain another reason for privileging it over Lagrangian mechanics.
@@
 Two photons with different momenta can follow the same trajectory in physical space.} Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
-Different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
+Moreover, different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
 In contrast, the trajectories in phase space (encoded by the Hamiltonian vector field) uniquely define the Hamiltonian up to a constant.} Hence, the set of classical systems that satisfy motion equivalence is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics.
@@
 
-Additionally, whereas North \parencites*[76]{North} seeks to determine the `real' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure ontologically, but as a formal device for encoding physical information or a calculational tool.
+Additionally, whereas North \parencites*[76]{North} \textbf{NOTE: before the citation was just \textbf{North} without the repetition} seeks to determine the `real' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure ontologically, but as a formal device for encoding physical information or a calculational tool.
 As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is `fundamental in the bottom-layer sense' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental in a classical world does not matter for the questions we seek to answer.
@@
 
-When physical quantities change continuously over time, $\mathcal{S}$ is a manifold, and we define $\mu(U) = \int_U \rho_c(s) d\mathcal{S} $.
+When the system is characterized by continuous quantities, $\mathcal{S}$ is a manifold, and we define $\mu(U) = \int_U \rho_c(s) d\mathcal{S} $.
 State variables $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates' and $\xi$ is a coordinate map.
@@
 Integrating over the closed surface of the parallelepiped, opposite sides provide equal and opposite contributions.
-Hence, the integral over this closed volume must equal zero.
+Hence, the integral over the closed surface surrounding this volume must equal zero.
 This shows that the 2-form $\omega$ must be closed.
@@
 \noindent
-\textit{Deterministic and reversible evolution}: given the present state of the system, all future (determinism) and past (reversible) states are uniquely identified. \es
+\textit{Deterministic and reversible evolution}: given the present state of the system, all future (determinism) and past (reversibility) states are uniquely identified. \es
 \end{quotation}
@@
 
-In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^i$.
-Therefore, for fixed $q$, the Jacobian matrix  $\frac{\partial v^i}{\partial p_j}$ must be invertible.
+In this case---which we call \textit{weak equivalence}---$v^i(q^j, p_k)$ is invertible at every $q^j$.
+Therefore, for fixed $q$, the Jacobian matrix  $\frac{\partial v^i}{\partial p_k}$ must be invertible.
 Let $H$ denote the Hamiltonian.
-By Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, so the Hessian $\frac{\partial^2 H}{\partial p_i \partial p_j} = \frac{\partial}{\partial p_i} (\frac{\partial H}{\partial p_j})$ equals the Jacobian $\frac{\partial v^i}{\partial p_j}$.
+By Hamilton's equations, $v^i = d_t q^i = \frac{\partial H}{\partial p_i}$, so the Hessian $\frac{\partial^2 H}{\partial p_k \partial p_i} = \frac{\partial}{\partial p_k} (\frac{\partial H}{\partial p_i})$ equals the Jacobian $\frac{\partial v^i}{\partial p_k}$.
 Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere---which we take to be positive.
@@
 \begin{equation} \label{massive_Hamiltonian}
-H = \frac{1}{2m} (p_j - A_j) g^{ij}(p_j - A_j) + V \\
+H = \frac{1}{2m} (p_i - A_i) g^{ij}(p_j - A_j) + V \\
 \end{equation}

### 2021-04-28 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory : first stab

\documentclass{article}

\usepackage{assumptionsofphysics}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\frenchspacing

\title{Bare minimum: set theory}

\date{\vspace{-5ex}}
\begin{document}

\maketitle

\begin{abstract}
This note presents a condensed summary of set theory.

Bare minima are not meant to be either didactic or complete. They are meant to condense areas of knowledge in a few pages and function as a knowledge check, refresher and/or reference. They collect what we feel is the bare minimum of a subject an intellectually curious mind should be aware of, particularly in the context of foundational questions. This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
\end{abstract}

\section{Introduction}

Set theory is important for those working on the Assumptions of Physics for at least three reasons. First, it is a foundational framework in mathematics. Second, it showcases a successful attempt at concept generalization. Third, it shows how a formal system is bootstrapped.

There are two rough branches of set theory: naive and axiomatic. Naive set theory is built on intuitive concepts, and as such is not fully formalized and is open to potential paradoxes. Axiomatic set theory provides a fully formalized axiomatic system that aims to close those problems. We will start with naive set theory, which is the best setting for defining all the basic notions and getting a sense of how the framework works. Then we will turn our attention to axiomatic set theory to get a sense of what problems it solves and how, and whether those solutions are aligned with the aims of the foundations of physics.

TODO: cite https://www.math24.net/topics-set-theory and set theory book

\section{Naive set theory}

We take sets and elements as primitive objects, with no formal definition. A \textbf{set} $A$ is a collection of elements. If an element $e$ belongs to the set we say $e$ \textbf{is in} $A$, noted $ e 
\in A$. A set $A$ can be defined by listing the elements, noted $A = \{ e_1, e_2, e_3 \}$. A set $A$ can be defined by specifying how to build the elements through a rule (i.e. predicate) $P(e)$, and is noted $A = \{e \, | \,  P(e) \}$. The rule may include symbols like for all $\forall$, exists $\exists$, logical connectors AND $\AND$, OR $\OR$ and $\NOT$.

\begin{defn}[Common sets] The \textbf{empty set} $\emptyset$ is the set with no elements. The \textbf{set of natural numbers} $\mathbb{N} = \{ 0, 1, 2, ... \}$. The \textbf{set of positive natural numbers} $\mathbb{N}^+ = \{n \in \mathbb{N} \, | \, n > 0\}$. The \textbf{set of integer numbers} $\mathbb{Z} = \{ ... , -2, -1, 0, +1, +2, ...\}$. The \textbf{set of rational numbers} $\mathbb{Q}$. The \textbf{set of real numbers} $\mathbb{R}$. The \textbf{set of complex numbers} $\mathbb{C}$.
\end{defn}

\subsection{Basic definitions}

\begin{defn}[Set relationships]
	Let $A$ and $B$ be two sets. If all elements in $A$ are also contained in $B$, we say $A$ is a \textbf{subset} of $B$, noted $A \subseteq B$. In this case, we also say $B$ is a \textbf{superset} of $A$, noted $B \supseteq A$. Additionally, if $B$ contains elements that $A$ does not contain, we say $A$ is a \textbf{strict subset} of $B$, noted $A \subset B$, and $B$ is a \textbf{strict superset} of $A$, noted $B \supset A$. If the two sets contain exactly the same elements, then we say the sets are \textbf{equal}, noted $A = B$. If the two sets have no element in common they are said to be \textbf{disjoint}.
\end{defn}

\begin{defn}[Power set]
	Given a set $A$, its \textbf{power set}, noted $\mathcal{P}(A)$ or $2^{A}$, is the set of all subsets of $A$. That is, $\mathcal{P}(A) = \{B \, | \, B \subset A \}$.
\end{defn}

\begin{defn}[Set operations]
	We define the following operations between two sets $A$ and $B$:
	\begin{description}
		\item[Union.] Noted $A \cup B$, the union of $A$ and $B$ is the set of all elements that are contained by $A$, $B$ or both.
		\item[Intersection.] Noted $A \cap B$, the intersection of $A$ and $B$ is the set of all elements that are contained by both.
		\item[Set difference (or relative complement).] Noted $A \setminus B$, $A$ minus $B$ is the set of all elements in $A$ that are not contained by $B$.
		\item[Complement (or absolute complement).] Noted $A^{\complement}$, represents all elements that are not contained in $A$. Note that his operation is context specific: we need to know that, within a certain context, $U$ is the set of all elements under study; then the complement of $A \subseteq U$ is $A^{\complement} = U \setminus A$.
	\end{description}
\end{defn}

We give the following informal definition. Given two elements $a$ and $b$, the \textbf{ordered pair} $(a, b)$ specifies two objects in that order. That is $(a, b) \neq (b, a)$ and $(a, a) \neq \{a\}$.

\begin{defn}[Cartesian product]
	Given two sets $A$ and $B$, the \textbf{Cartesian product} $A \times B = \{ (a,b) \, | \, \forall a \in A, b \in B \}$ is the set of all possible ordered pairs between the elements of $A$ and $B$.
\end{defn}

\subsection{Relations and function}

\begin{defn}[Binary relation]
	Given a set $A$, called \textbf{domain}, and a set $B$, called \textbf{codomain}, a \textbf{binary relation} is a set $R \subseteq A \times B$ of ordered pairs. We say $a$ is \textbf{$R$-related} to $b$, noted $aRb$ if $(a,b) \in R$. If $A=B$ the relation is said \textbf{homogeneous}, \textbf{heterogeneous} if not.
\end{defn}

\begin{defn}[Image and preimage]
	If $X \subseteq A$, the \textbf{image of $X$ under $R$} is the set of all elements in $B$ that are related to at least one element in $X$. The \textbf{image of $R$} is set of all elements in $B$ that are related to at least one element in $A$ (i.e. the image of the full set $A$).

	If $Y \subseteq B$, the \textbf{preimage of $Y$ under $R$} is the set of all elements in $A$ that are related to at least one element in $Y$. The \textbf{preimage of $R$} is set of all elements in $A$ that are related to at least one element in $B$ (i.e. the preimage of the full set $B$).
\end{defn}

\begin{defn}[Relation properties]
	Let $R$ be a binary relation between $A$ and $B$. We define the following properties:
	\begin{description}
		\item[Functional or right-unique.] An element $a \in A$ is related to at most one element $b \in B$. That is, if $aRb$ and $cRb$ then $a = c$.
		\item[Injective or left-unique.] An element $b \in B$ is related to at most one element $a \in A$. That is, if $aRb$ and $aRc$ then $b = c$.
		\item[Serial or left-total.] Every element $a \in A$ is related to at least one element $b \in B$. That is, for each $a \in A$ there exists at least one $b \in B$ such that $aRb$. In this case, the preimage of $R$ is the whole $A$.
		\item[Surjective or right-total.] Every element $b \in B$ is related to at least one element $a \in B$. That is, for each $b \in B$ there exists at least one $a \in A$ such that $aRb$. In this case, the image of $R$ is the whole $B$.
	\end{description}
	From those basic properties, we define the following:
	\begin{description}
		\item[One-to-one.] Injective and functional (i.e. left-unique and right-unique).
		\item[One-to-many.] Injective and not functional (i.e. left-unique and not right-unique).
		\item[Many-to-one.] Not injective and functional (i.e. not left-unique and right-unique).
		\item[Many-to-many.] Not injective nor functional (i.e. not left-unique nor right-unique).
	\end{description}	
\end{defn}

\begin{defn}[Function]
	A \textbf{partial function} $f : A \to B$ is a binary relationship between $A$ and $B$ that is functional (i.e. right-unique). The \textbf{graph} of the function $G \subset A \times B$ is the function expressed as ordered pairs. A \textbf{total function}, or simply a \textbf{function}, is a partial function that is also serial (i.e. left-total, defined on the whole domain).
\end{defn}

\begin{defn}[Composition]
	Let $A$, $B$ and $C$ be three sets. Let $R$ be a binary relationship between $A$ and $B$ and $S$ be a binary relationship between $B$ and $C$. Then the \textbf{composition of $R$ and $S$}, noted $S \circ R$, is the binary relationship between $A$ and $C$ such that $(a, c) \in S \circ R$ if we can find $b \in B$ such that $aRb$ and $bSc$.
\end{defn}

\begin{remark}
	The order of composition follows those of functions: if $f : A \to B$ and $g : B \to C$, composition leads to $c = g(f(a)) = (g \circ f)(a)$.
\end{remark}

\begin{prop}[Composition properties]
	Relation composition follows the following properties:
	\begin{itemize}
		\item Composition is associative: $R \circ (T \circ S) = (R \circ T) \circ S$
		\item Composition of functional/injective/serial/surjective relationship is functional/injective/serial/surjective.
		\item Specifically, if $f : A \to B$ and $g : B \to C$ are (partial) functions, then $g \circ f$ is a (partial) function.
	\end{itemize}
\end{prop}

\begin{defn}[Families]

\end{defn}

\begin{itemize}
	\item families (and operations on families)
	\item properties (boolean algebra, lattice)
	\item Ordered pairs and Cartesian product
	\item distribution rules of cartesian product
	\item power set
	\item cardinality
	\item Notable sets (empty, universal, natural, integers, rationals, reals, complex)
\end{itemize}

Functions
\begin{itemize}
	\item ...
\end{itemize}

\section{Axiomatic set theory}

Classes, elements, sets and proper classes.

%\bibliographystyle{plain}
%\bibliography{bibliography}

\end{document}

### 2021-04-28 Overviews/Informational granularity/InformationGranularity.tex
InfGran: removed most of the villegas (no time to review properly)

@@
 
-\usepackage{assumptionsofphysics}
+\usepackage[sectionnum]{assumptionsofphysics}
 
@@
 
-\def\eqgran{\doteq}
-\def\finer{\leqdot}
-\def\nfiner{\nleqdot}
-\def\coarser{\geqdot}
-\def\ncoarser{\ngeqdot}
-\def\sfiner{\lessdot}
-\def\scoarser{\gtrdot}
-
 \section{Introduction}
@@
 We were not able to find similar work in the literature.\footnote{If the reader has relevant suggestions, we would appreciate them.} The closest is in the context of comparative probability\cite{definetti, villegas} from which we have taken some mathematical ideas.
-The problem is that the preorder they typically use is essentially equivalent to the probability measure they recover, while the whole point in our case, as mentioned in the introduction, is that the preorder should give us a richer structure.
+The survey \cite{fishburnsurvey} contains some examples of different axioms which may be very helpful.
+In general, the problem is that the preorder they typically use is essentially equivalent to the probability measure they recover, while the whole point in our case, as mentioned in the introduction, is that the preorder should give us a richer structure.
 
@@
 Then $\stmt_1$ is \textbf{smaller than} $\stmt_2$ \textbf{in} $\stmt[u]$, noted $\stmt_1 \lequ \stmt_2$, if $\stmt_1 \finer \stmt_2$ or $\stmt_1 \equ \stmt_2$.
-We define $\stmt_1\leu\stmt_2$ by $\stmt_1\lequ\stmt_2$ and not $\stmt_1\equ\stmt_2$.
+We define $\stmt_1\ltu\stmt_2$ by $\stmt_1\lequ\stmt_2$ and not $\stmt_1\equ\stmt_2$.
 \end{defn}
@@
 \begin{prop}
-	If for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$, then we have $\stmt_1 \leu \stmt_2$. 
+	If for every statement $\stmt[v]$ infinitesimal in $\stmt[u]$, we have $\stmt_1\OR\stmt[v] \sfiner \stmt_2$, then we have $\stmt_1 \ltu \stmt_2$. 
 \end{prop}
@@
 
+A hint in that direction is given by the following proposition, given with no proof by \cite{villegas}.
+
+\begin{prop}
+	A necessary and sufficient condition for $\lequ$ to be monotonely continuous is the following.
+Let  $\{\stmt_i\}_{i=1}^{\infty}$ be a monotone sequence, that is $\stmt_i \narrower \stmt_j$ for all $i \leq j$ and $\stmt = \bigOR_i\stmt_i$ the disjunction, and let $\stmt[t] \in \tdomain$ such that $\stmt[t] \ltu \stmt$, then there exists an $n > 0$ such that $\stmt[t] \ltu \stmt_n$.
+\end{prop}
+
 Before achieving the above, it may be possible to obtain a related result about fineness and monotone continuity, which could translate to monotone continuity for $\lequ$.
@@
 
-Tentative layout:
-- summarize the result in terms of input requirements (stated with our notation) and output
-- discussion of how far we are from getting those requirements
-- sketch of the main steps of the proof (the key propositions and the approach of the proof of each in our notation)
-- discussion on whether it is feasable and how to relax the proof for finite additivity
-
-The paper \cite{villegas} by Villegas gives what seems to be an extremely relevant overview of qualitative probability, including many of the axioms which we seek in our structures.
-
-In this part, we will assume that $\tdomain_{\stmt[u]}$ is a subset of the theoretical domain on which $\lequ$ is total, $\stmt[u]\in \tdomain_{\stmt[u]}$ is a distinguished element (we'll refer to it as the \emph{unit} when needed).
-We will use the word ``equiprobable" for $=_{\stmt[u]}$ since it means something different from equigranular (since $\leq_{\stmt[u]}$ is a different ordering from fineness, even though it will be derived from fineness).
-
-The goal is to use $\lequ$ as a qualitative probability.
-The first basic requirement is that $\leq_{\stmt[u]}$ is a total pre-order with maximal and minimal elements.
-These were contained in several desiderata earlier.
-
-Below is the version of the Axiom of Monotony for $\lequ$; earlier in Proposition \ref{fineness_monotony} it was shown that fineness satisfies this, and Definition \ref{smalleru} would allow this to descend to $\lequ$ as well.
-
-Axiom of Monotony: 
-Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
-Assume that $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, and $\stmt_2\ncomp \stmt_4$ (i.e.
-$\stmt_2\AND \stmt_4 =_{\stmt[u]} \impossibility$).
-Then we have $\stmt_1\OR \stmt_3 \leq_{\stmt[u]} \stmt_2\OR \stmt_4.$
-
-The following are consequences of $\lequ$ being a total pre-order with the Axiom of Monotony. Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
+The paper \cite{villegas} by Villegas gives us some conditions under which we can recover a probability measure given a preorder.
+Here we state the conditions required in terms of $\lequ$ and $\tdomain_{\stmt[u]}$, which we take it to be the set of statements narrower than $\stmt[u]$.
+We start from the axioms of pre-order:
+
 \begin{itemize}
-    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ and $\stmt_2<_{\stmt[u]} \stmt_3$, then $\stmt_1<_{\stmt[u]}\stmt_3$. 
-    \item If $\stmt_1\narrower \stmt_2$ then $\stmt_1\leq_{\stmt[u]} \stmt_2$
-    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, $\stmt_2\leq_{\stmt[u]} \stmt_4$, and $v\leq_{\stmt[u]} \stmt_1$, then $\stmt_2 \AND \NOT \stmt_1 \leq_{\stmt[u]} \stmt_4 \AND \NOT \stmt_3$
-    \item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ then $\NOT \stmt_2 \leq_{\stmt[u]}\NOT \stmt_1$
+	\item Totality: If $\stmt_1,\stmt_2\in \tdomain_{\stmt[u]}$, then $\stmt_1 \lequ \stmt_2$ or $\stmt_2 \lequ \stmt_1$
+	\item Reflexivity: For all statements $\stmt\in \tdomain_{\stmt[u]}$, we have $\stmt\lequ\stmt$
+	\item Transitivity: For all statements $\stmt_1,\stmt_2,\stmt_3\in \tdomain_{\stmt[u]}$, if $\stmt_1\lequ\stmt_2$ and $\stmt_2\lequ\stmt_3$, then $\stmt_1\lequ\stmt_3$.
+	\item Boundedness: We have $\impossibility\ltu\certainty$. For all $\stmt\in \tdomain_{\stmt[u]}$ we have $\impossibility\leq_{\stmt[u]}\stmt\lequ\certainty$
 \end{itemize}
 
-Beyond the above, we will also utilize Monotone Continuity (Definition \ref{monotone_continuity}) in much of what follows.
-
-Another important notion for qualitative probability is the idea of atoms.
+Next is the axiom of monotony.
+Let $\stmt_1,\stmt_2,\stmt[t]_1,\stmt[t]_2 \in \tdomain_{\stmt[u]}$ with $\stmt[t]_1 \ncomp \stmt[t]_2$.
+Then from $\stmt_1 \lequ \stmt[t]_1$ and $\stmt_2 \lequ \stmt[t]_2$ if follows that $\stmt_1\OR \stmt_2 \lequ \stmt[t]_1\OR \stmt[t]_2$.
+Moreover, if $\lequ$ is replace by $\ltu$ in one of the first two equality, then the last one holds with $\ltu$.
+
+Apart from totality, we have similar properties already defined on fineness.
+We will need to understand on what conditions they can be carried to $\lequ$.
+The best way forward will be to assume totality and see whether the other properties can be derived at least in that case.
+
+Next we have the axiom of monotone continuity (Definition \ref{monotone_continuity}). This we have seen before and we suspect is connected to the lack of countable infinitesimals.
+
+The last condition is the lack of atoms. An atom is defined in the following way:
 
 \begin{defn}
-An \textbf{atom} is a statement $\stmt\in S$ such that $\impossibility<_{\stmt[u]}\stmt$ and there does not exist any statement $\stmt_2\in S$ such that $\impossibility<_{\stmt[u]} \stmt_2<_{\stmt[u]} \stmt$.
-\end{defn}
-In our framework, if possibilities are equigranular and not equivalent at $\stmt[u]$ scale to $\impossibility$, then they will be atoms.
+An \textbf{atom} is a statement $\stmt\in S$ such that $\impossibility \ltu \stmt$ and there does not exist any statement $\stmt_2\in S$ such that $\impossibility \ltu \stmt_2 \ltu \stmt$. 
+\end{defn}
+
+Essentially, an atom is an immediate successor of the impossibility.
+In our framework, the possibilities are the immediate successor of the impossibility under narrowness and fineness.
+If the unit contains finitely many possibilities, then we will have atoms for $\lequ$.
+However, in that case we will assume equigranular possibilities, and therefore the counting measure will apply.
+If the unit has infinitely many possibilities and the possibilities are equigranular, then all possibilities are infinitesimals.
+We should be able to show that, in that case, there are no atoms.
+
 The presence or lack of atoms has several important consequences.
@@
 The survey \cite{fishburnsurvey} contains some examples of different axioms which may be very helpful.
-
-\subsection{Old notes}
-
-\section{Adapting results from \cite{villegas}}
-
-In this part, we will assume that $\tdomain_{\stmt[u]}$ is a subset of the theoretical domain, $\stmt[u]\in \tdomain_{\stmt[u]}$ is a distinguished element (we'll refer to it as the \emph{unit} when needed), and there is a total pre-order $\leq_{\stmt[u]}$ defined on $\tdomain_{\stmt[u]}$.
-We will write $=_{\stmt[u]}$ for $\leq_{\stmt[u]}$ and $\geq_{\stmt[u]}$, and we write $<_{\stmt[u]}$ for $\leq_{\stmt[u]}$ and not $=_{\stmt[u]}$.
-We will use the word ``equiprobable" for $=_{\stmt[u]}$ since it means something different from equigranular (since $\leq_{\stmt[u]}$ is a different ordering from fineness, even though it will be derived from fineness).
-We assume $\tdomain_{\stmt[u]}$ is closed under complement and countable union (and hence countable intersection), and that there are two additional elements, $\certainty$ and $\impossibility$, which are maximal and minimal in this ordering, respectively.
-
-\begin{remark}
-	In defining the subset $\tdomain_{\stmt[u]}$ and the total preorder $\leq_{\stmt[u]}$, we will have to be very careful.
-We want to ``squash" together all the sets which should be measure zero in this unit (e.g.
-anything ``lower-dimensional"), ``blow up" all the ``infinite measure" subsets (e.g.
-``higher dimensional"), and ensure that everything is compatible under this order, so there can be only one ``type" of property under consideration.
-\end{remark}
-
-We write out the main axioms and statements from \cite{villegas} which interest us in this situation, in the notation established earlier. On page 1789, we have the following:
-
-Axiom of Pre-order:
-\begin{itemize}
-	\item Totality: If $\stmt_1,\stmt_2\in \tdomain_{\stmt[u]}$, then $\stmt_1 \leq_{\stmt[u]} \stmt_2$ or $\stmt_2\leq_{\stmt[u]} \stmt_1$
-	\item Reflexivity: For all statements $\stmt\in \tdomain_{\stmt[u]}$, we have $\stmt\leq_{\stmt[u]}\stmt$
-	\item Transitivity: For all statements $\stmt_1,\stmt_2,\stmt_3\in \tdomain_{\stmt[u]}$, if $\stmt_1\leq_{\stmt[u]}\stmt_2$ and $\stmt_2\leq_{\stmt[u]}\stmt_3$, then $\stmt_1\leq_{\stmt[u]}\stmt_3$.
-	\item Existence of max and min: We have $\impossibility\leq_{\stmt[u]}\certainty$ and \emph{not} $\certainty\leq_{\stmt[u]}\impossibility$.
-For all $\stmt\in \tdomain_{\stmt[u]}$ we have $\impossibility\leq_{\stmt[u]}\stmt\leq_{\stmt[u]}\certainty$
-\end{itemize}
-
-The above list is simply the requirement that $\leq_{\stmt[u]}$ is a total pre-order and that there are a maximal and minimal element.
-
-Axiom of Monotony: 
-Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
-Assume that $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, and $\stmt_2\ncomp \stmt_4$ (i.e.
-$\stmt_2\AND \stmt_4 =_{\stmt[u]} \impossibility$).
-Then we have $\stmt_1\OR \stmt_3 \leq_{\stmt[u]} \stmt_2\OR \stmt_4.$
-
-Monotony guarantees that given two incompatible statements, each ``bigger" than another statement, then their disjunction is still bigger than the disjunction of the smaller statements.
-
-The following are consequences of the two axioms above. Let $\stmt_1,\stmt_2,\stmt_3,\stmt_4\in \tdomain_{\stmt[u]}$.
-\begin{itemize}
-	\item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ and $\stmt_2<_{\stmt[u]} \stmt_3$, then $\stmt_1<_{\stmt[u]}\stmt_3$. 
-	\item If $\stmt_1\narrower \stmt_2$ then $\stmt_1\leq_{\stmt[u]} \stmt_2$
-	\item If $\stmt_1\leq_{\stmt[u]} \stmt_2$, $\stmt_3\leq_{\stmt[u]} \stmt_4$, $\stmt_2\leq_{\stmt[u]} \stmt_4$, and $v\leq_{\stmt[u]} \stmt_1$, then $\stmt_2 \AND \NOT \stmt_1 \leq_{\stmt[u]} \stmt_4 \AND \NOT \stmt_3$
-	\item If $\stmt_1\leq_{\stmt[u]} \stmt_2$ then $\NOT \stmt_2 \leq_{\stmt[u]}\NOT \stmt_1$
-\end{itemize}
-
-Next, we will need a more subtle axiom in order to make sense of infinite combinations of statements.
-
-\begin{remark}
-	The \emph{fineness} relation $\finer$ should satisfy all of the axioms of pre-order and the axiom of monotony, except for totality. We restrict to $\tdomain_{\stmt[u]}$ in order to get totality. 
-\end{remark}
-
-Axiom of Monotone Continuity:
-Let $\{\stmt_i : i\in \mathbb{Z}^+\}$ be a sequence of statements which is monotone increasing with respect to narrowness (containment), and let $\stmt = \cup_i\stmt_i$ be the disjunction.
-Let $t\in \tdomain_{\stmt[u]}$ such that $\stmt_i\leq_{\stmt[u]} t$ for all $i$.
-Then $\stmt\leq_{\stmt[u]} t$.
-
-In other words, an infinite disjunction can only strictly exceed a statement in coarseness if the disjunction of a finite subset of them already exceeds it.
-Here is an interesting consequence of the combination of the above axioms:
-
-\begin{prop}[Corollary to Lemma 1 in \cite{villegas}]
-	\label{infinitenull}
-	Let $\{\stmt_i : i\in \mathbb{Z}^+\}$ be an infinite collection such that $\stmt_i\AND\stmt_j =_{\stmt[u]} \impossibility$ for all $i\neq j$, and $\stmt_i =_{\stmt[u]} \stmt_j$ for all $i,j$.
-Then it must be the case that $\stmt_i=_{\stmt[u]}\impossibility$ for all $i$.
-\end{prop}
-This says that infinite collections of equiprobable statements with null overlaps must be null themselves.
-
-An \emph{atom} is a statement $\stmt\in S$ such that $\impossibility<_{\stmt[u]}\stmt$ and there does not exist any statement $\stmt_2\in S$ such that $\impossibility<_{\stmt[u]} \stmt_2<_{\stmt[u]} \stmt$.
-In particular, Theorem 5 in \cite{villegas} \S3 includes the fact that $\tdomain_{\stmt[u]}$ is atomless if and only if every statement can be partitioned into two equiprobable statements.
-Theorem 3 in \S4 states that if $\tdomain_{\stmt[u]}$ is atomless, then there exists a unique probability measure compatible with $\leq_{\stmt[u]}$, and it is countably additive!
-
-\subsection{Using Monotone Continuity}
-
-The core of the proof of Theorem 3 in Section 4 of \cite{villegas} comes from Theorem 5 (iv) in Section 3.
-In the terminology of Villegas, a \emph{qualitative probability $\sigma$-algebra} is a $\sigma$-algebra of subsets of a space with an ordering that satisfies the Axiom of Pre-order, the Axiom of Monotony, and the Axiom of Monotone Continuity.
-We also need the language of \emph{atoms}, defined earlier.
-We will reproduce some of the ideas in the proofs of these theorems in the language of statements and our theoretical domain, while attempting to keep track of the usages of Monotone Continuity.
-The hope is that there is a weakened version of the axiom that gets us the desired result, or that we can build the ordering $\leq_{\stmt[u]}$ with Monotone Continuity but it can still distinguish the infinite sets described earlier.
-
-Let $A$ be a statement in $\tdomain_{\stmt[u]}$ with $A>_{\stmt[u]} \impossibility$. A \emph{minor incomplete partition} of $A$ is a collection $M_A$ of statements with the following properties:
-\begin{itemize}
-	\item We have $\stmt>_{\stmt[u]}\impossibility$ and $\stmt \narrower A$ for all $\stmt\in M_A$, that is, they are non-null and narrower than $A$
-	\item For all $\stmt_1,\stmt_2\in M_A$, we have $\stmt_1\AND\stmt_2 =_{\stmt[u]} \impossibility$, that is, they overlap only on statements ``squashed" to zero. 
-	\item The disjunction of $M_A$ is finer than its complement in $A$: $$\bigOR_{\stmt\in M_A}\stmt \finer (A\AND \NOT\bigOR_{\stmt\in M_A}\stmt).$$ This is the condition of being \emph{minor}.
-	%UM_A finer than A - UM_A^C
-\end{itemize}
-
-It is shown in \cite{villegas} that any (minor) incomplete partition is at most countable (Theorem 3).
-Rather than give the full proof of this fact, we note that it follows from a large collection of lemmas, the most interesting of which is the following, which we will prove below:
-
-\begin{prop}[Lemma 1 from \cite{villegas}]
-	If $\{\stmt[a]_i: i\in I\}$ is an infinite set of statements such that (1) $\stmt[a]_i \AND \stmt[a]_j =_{\stmt[u]}\impossibility$ for all $i\neq j$ and (2) $\stmt[a]_i>_{\stmt[u]}\impossibility$ for all $i$, then the following holds.
-Given a statement $\stmt>_{\stmt[u]}\impossibility$, only finitely many of the $\stmt[a]_i$'s have $\stmt[a]_i>_{\stmt[u]}\stmt$.
-\end{prop}
-\begin{proof}
-	By way of contradiction, let $\stmt[a]_1,\ldots$ be a sequence of statements violating the conclusion (i.e.
-infinitely many statements $\stmt[a]_i$ with $\stmt[a]_i>_{\stmt[u]}\stmt>_{\stmt[u]}\impossibility$.
-Consider the new sequence $$\stmt_i = \bigOR\limits_{n=i}^{\infty}\stmt[a]_n.$$ This sequence is monotonely decreasing to the impossibility in narrowness (and hence in fineness and in $>_{\stmt[u]}$), but we also have $\stmt_i\geq_{\stmt[u]}\stmt$ for all $i$.
-By monotone continuity (the contrapositive), we have $\stmt[s]\leq_{\stmt[u]}\impossibility$, a contradiction.
-\end{proof}
-
-A quick corollary of this is Proposition \ref{infinitenull}, above. Using minor incomplete partitions, we prove the following.
-
-\begin{prop}[Theorem 4, \S3 \cite{villegas}]
-	\label{split}
-	If there are no atoms, then for every statement $\stmt$ there exist statements $\stmt_1$ and $\stmt_2$ such that: $\stmt_1\ncomp \stmt_2$, $\stmt_1\equ\stmt_2$, and $\stmt_1\OR\stmt_2 = \stmt$.
-In other words, $\stmt$ can be partitioned into two equiprobable (note: I remember we don't like this word but I don't remember if we replaced it or not) statements.
-\end{prop}
-\begin{proof}
-	Let $S$ be the set of minor incomplete partitions of $\stmt$, and endow it with a partial order by inclusion.
-Given a totally ordered chain in $S$, its union is still a minor incomplete partition, and is thus an upper bound for the chain.
-We may apply Zorn's lemma to find a maximal element of $S$, the disjunction of which we call $\stmt_1$.
-We claim that $\stmt_1$ and $\stmt_2:= \stmt\AND\NOT\stmt_1$ are equiprobable, incompatible statements.
-Incompatibility is immediate from the definition.
-For equiprobability, first note that by definition of being a minor incomplete partition we have $\stmt_1\lequ\stmt_2$.
-Next, if $\stmt[t]_1,\ldots$ is a monotone decreasing (with respect to $\lequ$) sequence of statements $\stmt[t]_n>_{\stmt[u]}\impossibility$ narrower than $\stmt_2$ and converging to a zero-comparable event, we have that $\stmt_1\OR\stmt[t]_n>_{\stmt[u]}\stmt_2\OR\stmt[t]_n$ since $\stmt_1$ is maximal.
-We needed no atoms in order to guarantee this sequence exists (Lemma 3 in Section 3 of \cite{villegas}).
-It follows by monotone continuity that $\stmt_1\geq_{\stmt[u]} \stmt_2$.
-\end{proof}
-
-We will also need a few types of \emph{random variables} as defined in \cite{villegas} \S3.
-Let $A({\stmt[u]}) = \{x \in X : x \comp \stmt[u]\}$ be the set of possibilities associated with $\stmt[u]$.
-This is a subset of the collection of possibilities which will serve as the ``sample space" after picking a unit $\stmt[u]$.
-
-A \emph{uniform sequence of finitely-valued random variables} is a sequence of functions (random variables in the traditional sense) $X_1,X_2,\ldots: A({\stmt[u]})\to E$ where $E$ is a finite set, with the following property: for all $\{x_1,\ldots,x_n\}\subseteq E$, all the events $$\{p\in A({\stmt[u]}): X_i(p) = x_i, i=1,\ldots,n\}$$
-are equiprobable ($\equ$ with each other).
-
-Notice that monotone continuity allows us to let $n\to\infty$ to define events where the whole sequence is specified with the resulting events all being equiprobable.
-By Proposition \ref{infinitenull} above, all such events are null.
-
-Building on these sequences of random variables, we will use \emph{uniform random variables} to define probability measures.
-Let $X:A({\stmt[u]})\to[0,1]$ be a random variable, and let $I$ and $J$ be intervals contained in $[0,1]$.
-We say $X$ is a uniform random variable if $X^{-1}(I)\lequ X^{-1}(J)$ if and only if $|I|\leq|J|$ for all such $I,J$.
-
-From Proposition \ref{infinitenull}, $X^{-1}(a)$ is a zero-comparable event for all $a\in[0,1]$, and if $0=a_0<a_1<\cdots<a_n=1$, then $X^{-1}([a_i,a_{i+1}))$ are a finite partition of $\tdomain$.
-
-Next, we will pull many of the above concepts together to show how we obtain a probability measure on $A({\stmt[u]})$ realized as a function on $\tdomain$ compatible with $\lequ$.
-
-\begin{prop}\label{uniformexist}
-	If $\tdomain_{\stmt[u]}$ has no atoms with respect to $\lequ$, then there exists a uniform random variable.
-\end{prop}
-\begin{proof}
-	By Proposition \ref{split} we may split any statement into two equiprobable, null-compatible sets.
-Starting with the statement $\stmt[u]$ (here, the ``universe" which we want to have probability 1), repeatedly divide into two statements in stages.
-At stage $n$ (starting with $n=0$), we will have statements $\stmt[u]^{(n)}_1,\ldots,\stmt[u]^{(n)}_{2^n}$.
-Now, take the disjunction of alternating statements in each collection so that there is a ``right" and ``left" (odd and even index) half at each stage.
-For example, at stage $n=1$, the odd and even halves are simply $\stmt[u]^{(1)}_1$ and $\stmt[u]^{(1)}_2$; at stage $n=2$ we have $\stmt[u]^{(2)}_1\OR\stmt[u]^{(2)}_3$ and $\stmt[u]^{(2)}_2\OR\stmt[u]^{(2)}_4$.
-
-	Define a sequence of random variables $X_n:A(\stmt[u])\to\{0,1\}$ by $X_n(a) = 1$ if and only if $a$ is in the odd half in stage $n$.
-This gives each $a\in A(\stmt[u])$ a binary ``address".
-By the construction, it follows that this is a uniform sequence of finitely-valued random variables (defined above).
-
-	Now, we define a uniform random variable $X:A(\stmt[u])\to[0,1]$.
-For $a\in A(\stmt[u])$, define $X(a) = \sum_{n=1}^{\infty}2^{-n}X_n(a)$.
-In other words, $X(a)$ is the binary expansion $0.X_1(a)X_2(a)\ldots_2$.
-
-	One can show immediately that $X^{-1}(I)\lequ X^{-1}(J)$ if and only if $|I|\leq |J|$ when $I$ and $J$ have endpoints on numbers with finite binary expansions (since $X_1,\ldots,X_n$ are uniform), and by monotone continuity, the uniformity condition extends to arbitrary intervals in $[0,1]$.
-\end{proof}
-
-Finally, we reach our goal of building a probability measure on $A(\stmt[u])$.
-
-\begin{prop}
-	If $\tdomain_{\stmt[u]}$ has no atoms with respect to $\lequ$, then there exists a unique compatible probability measure, and it is countably additive. 
-\end{prop}
-\begin{proof}
-	By Propsition \ref{uniformexist}, let $X:POSS_{\stmt[u]}\to[0,1]$ be a uniform random variable.
-	%Define $F:[0,1]\to\tdomain$ by $F(x) = X^{-1}([0,x])$.
-It is straightforward to see that $F$ is monotone in $x$ (with respect to $\lequ$), and that if $x_n\to x$, then $\lim_nF(x_n) \equ F(x)$.
-%Started writing this part from paper, but it is not directly used in construction.
-Will add details later as needed.
-	We claim that for any statement $\stmt\in\tdomain_{\stmt[u]}$, there exists a unique number $p\in[0,1]$ such that $\stmt\equ X^{-1}([0,p])$.
-Define $P:\tdomain_{\stmt[u]}\to[0,1]$ by $P(\stmt)= p$.
-This defines a probability measure compatible with $\lequ$.
-Details forthcoming as needed....
-\end{proof}
-
-Ideas involved: Defines a function $X:A(\stmt[u])\to [0,1]$ in a way that is consistent with $\lequ$ by having ``larger" sets of possibilities (i.e.
-coarser statements) corresponding to larger intervals in $[0,1]$.
-Then, the measure of any other statement is given by finding an equiprobable statement which is the preimage of $[0,p]$ under $X$ and defining the measure to be $p$.
-
-Wish: find a way to prove the above statements in an order-theoretic way.
-In particular, we wish to show that the order type of the equigranular-equivalence classes is equivalent to (a subset of) $\mathbb{R}$.
-Sufficient conditions to be isomorphic to $\mathbb{R}$ are (1) lacking endpoints (2) complete (i.e.
-has Least Upper Bound property) (3) has dense countable subset.
-The use of random variables may be very similar to an approach using order theory.
-For example, the transition from a uniform sequence of finitely-valued random variables into a (single) uniform random variable makes use of a countable dense subset.
-
-% END OF NOTES
 

### 2021-04-30 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
ClassMechPhil: finished final review

@@
 By representing the distribution, the density quantifies the amount of stuff in each phase space state.
-Typically, the canonical position and momentum observables play the role of specifying a classical state, since they fix the values of all other observables.
+Typically, the canonical position and momentum play the role of specifying a classical state, since they fix the values of all other observables.
 Ruetsche refers to this as ``fundamental in the physicist's sense'' \parencite*[31, 200]{Ruetsche}.
@@
 In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
-If the value of the density at one point is larger than another, this entails that the density is higher at that point of phase space.
-Physically, this means that there is more matter at this point.
+If the value of the density at one point is larger than another, physically it means that there is more matter at this point.
 Likewise, if the value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
-For brevity, we'll refer to this as the property of \textit{point-wise comparison}.
+For brevity, we'll refer to this as allowing \textit{point-wise comparison}.
 Point-wise comparison requires a coordinate-invariant density.
 As the derivation in Section~\ref{infinitesimal} shows, to guarantee the existence of an invariant density, we must define state space over a symplectic manifold.
-In other words, to generically have a density with the property of point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
+In other words, to generically have a density which allows point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
 
@@
 The result is that we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in (in the sense of space-time coordinates).
-Any and all (time-independent) canonical coordinate systems will support the formal property of point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space $(q^i, p_i, t, E)$.
+Any and all (time-independent) canonical coordinate systems will support point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space $(q^i, p_i, t, E)$.
 See \textcites[]{Struckmeier}.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.\footnote{One might wonder if it is possible to construct a coordinate-independent density, perhaps using either (i) volume forms from differential geometry or (ii) the canonical volume form of Riemannian geometry.
@@
 Hence, we write $-\vector{\theta} = (-p, 0, H) $ as the vector potential for the displacement vector $\vector{S} $.
-By matching components of $\vector{S}$ and $-curl(\theta)$, we show that $H$ is the Hamiltonian for the system.
+By matching components of $\vector{S}$ and $-curl(\vector{\theta})$, we show that $H$ is the Hamiltonian for the system.
 First, by definition, $S_q \equiv \frac{dq}{dt} $ and $S_p \equiv  \frac{dp}{dt} $.
@@
 Since the trajectories $\gamma$ and $\gamma'$ have the same endpoints, they define a closed-curve $C$, enclosing a region $\Sigma $.
-Hence, by Stoke's theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot dC = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ .
+Hence, by Stokes' theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot \vector{dC} = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ .
 Recalling that i) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and ii)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $curl(\vector{\theta})$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
-This establishes a more general Hamilton's principle of stationary action, which we have derived simply by assuming that the displacement vector is divergenceless.
+This establishes a more general version of Hamilton's principle of stationary action, which we have derived simply by assuming that the displacement vector is divergenceless.
 In the special case that motion equivalence holds, the Legendre transformation is well-defined and $L_p = p \dot{q} - H(q, p, t)$ equals the usual Lagrangian $L(q, \dot{q}, t)$, thereby recovering the principle of stationary action for Lagrangian mechanics.
@@
 Additionally, our definition accommodates models of classical massless particles by maintaining the uniqueness of solutions, even outside the hyper-regular domain.
-Moving beyond classical mechanics, our framework has the advantage of providing an elegant relationship with quantum mechanics.
-By changing the requirement of infinitesimal reducibility to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.\footnote{We provide details of the quantum case elsewhere.} Hence, our framework perspicuously distinguishes quantum from classical systems.
+Moving beyond classical mechanics, our framework has the advantage of providing an elegant relationship with both classical statistical mechanics and quantum mechanics.
+We note that the interpretation of the symplectic form is exactly the one used by the fundamental postulate of statistical mechanics, which uses phase space volumes (i.e.
+the volume form generated by the symplectic form) to count states.
+Additionally, by changing the requirement of infinitesimal reducibility to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.\footnote{We provide details of the quantum case elsewhere.} Hence, our framework perspicuously distinguishes quantum from classical systems.
 We take these considerations as reason enough for adopting our preferred notion of a classical system.

### 2021-04-30 Overviews/Informational granularity/InformationGranularity.tex
InfGran: draft ready

@@
 The objective is to find necessary and/or sufficient conditions for such constructions which can be given a straightforward physical motivation.
-%    As in comparative probability, we start with a $\sigma$-complete Boolean algebra that represents all possible descriptions in a given context and impose a preorder.
-In our case, the preorder does not represent likelyhood comparison, but rather whether one description is at a finer level of granularity than another.
-At this point, any statement (apart from an impossible one) can be chosen to act as a unit, and a measure (not necessarily bounded) can be constructed to quantify the degree to which the other statements give a more or less refined account.
-The starting lattice is not necessarily Archimedean, therefore the choice of unit will decide which elements are finite, infinite or infinitesimal.
-The family of measures so obtained are able to give a size to sets at different level of scale (e.g.
-points, lines, surfaces, ...) much like geometrical structures do.
-Our hope is that this structure recovers as special cases the geometrical, measure theoretical and probabilistic constructions we currently use in the sciences while giving a generalization that still works in the case where they fail (e.g.
-lack of $\sigma$-additivity in quantum probability) or expected to fail (e.g.
-breakdown at Planck's length).
 
@@
 The overall goal of Assumptions of Physics is to identify a handful of physical principles from which the basic laws can be rigorously derived.
-We have already established that topologies and $\sigma$-algebras can be recovered from requirements of experimental verifiability.\cite{aop-book,aop-overview-verifiability} See also (TODO: Kevin Kelly).
+We have already established that topologies and $\sigma$-algebras can be recovered from requirements of experimental verifiability.\cite{aop-book,aop-overview-verifiability,aop-math-topologydistinguishability} For similar ideas see also \cite{kelly1996logic}.
 The next step is to understand the underpinnings of geometric and measure theoretic structures (including probability) which are layered on top of topologies and $\sigma$-algebras.
 As usual, a formal construction from scratch will force us to make explicit the physical requirements that are baked into the use of the common concepts and provide us with a framework that is, hopefully, still valid when those assumptions fail.
-%TODO: add topology proceeding, update book reference
 
@@
 \textbf{The working hypothesis is that this, and only this, is what geometric and measure theoretic structures capture: how coarse or fine grained the statements are within a scientific theory.} Finding the fundamental axioms associated to the notion of information granularity is therefore the central problem.
-
-%This working draft summarizes the status of our work on a new approach to the foundations of geometry and measure theory and is intended to gather early feedback.
-The goal is to rederive the usual concepts in geometry, measure theory and probability theory from a simpler more general structure, namely a preorder.
-This construction will
-
-%The general idea is similar to what is done in comparative probability, yet we want the techniques extend from a single measure to a family of measures that allow us to define and compare sizes of sets at different scales (e.g.
-points, lines, surfaces, ...).
 
@@
 \begin{defn}
-	Let $\stmt_1, \stmt_2 \in \tdomain$ be two statements such that $\stmt_1 \finer \stmt_2$.
-We say that they are \textbf{finitely comparable} to each other if we can find a finite set $\{ \stmt[t]_i \}_{i=1}^n$ such that $\stmt[t]_i \finer \stmt_1$ for all $i \leq n$ and $\stmt_2 \finer \bigOR_{i=1}^n \stmt[t]_i$.
-\end{defn}
-
+	Let $\stmt, \stmt[u] \in \tdomain$ be two statements such that $\stmt \finer \stmt[u]$.
+We say that they are \textbf{finitely comparable} to each other if we can find a finite set $\{ \stmt_i \}_{i=1}^n$ such that $\stmt_i \finer \stmt$ for all $i \leq n$ and $\stmt[u] \finer \bigOR_{i=1}^n \stmt_i$.
+\end{defn}
+
+TODO[Mark]: check the proof
 \begin{prop}
@@
 \begin{proof}
-	TODO
+	Suppose $\stmt$ and $\stmt[u]$ are finitely comparable.
+Then we can find a set of $n$ statements $\{ \stmt_i \}_{i=1}^n$ such that $\stmt_i \finer \stmt$ for all $i \leq n$ and $\stmt[u] \finer \bigOR_{i=1}^n \stmt_i$.
+Now suppose we have a set of statements $\{ \stmt[t]_i \}_{i=1}^n$ such that $\stmt \finer \stmt[t]_i$ for all $i$ and $\stmt[t]_i \ncomp \stmt[t]_j$ for all $i \neq j$.
+Then we also have $\stmt_i \finer \stmt \finer \stmt[t]_i$ for all $i$.
+By \ref{fineness_monotony} we have $\bigOR_{i=1}^n \stmt_i \finer \bigOR_{i=1}^n \stmt[t]_i$ and therefore $\stmt[u] \finer \bigOR_{i=1}^n \stmt_i \finer \bigOR_{i=1}^n \stmt[t]_i$.
+Therefore we cannot find a set of $n$ statement that satisfy the definition of infinitesimal and therefore $\stmt$ is not an infinitesimal with respect to $\stmt[u]$.
 \end{proof}
@@
 \begin{proof}
-	TODO
+	Since $\impossibility \finer \stmt$ for any $\stmt$, we have $\impossibility\lequ\stmt$ regardless of the choice of $\stmt[u]$.
+Conversely, since $\stmt \finer \certainty$ for any $\stmt$, we have $\stmt\lequ\certainty$ regardless of the choice of $\stmt[u]$.
 \end{proof}
@@
 We will need an interesting-enough subset of $\tdomain$ to capture the desired notions while still having $\lequ$ as a total order.
-
-% case: possibilities in u: (1) finite: impose counting measure (2) countable: finitely additive but not countably (3) uncountable: countably additive, legitimate measure
 
@@
 
-\subsection{Distribution on natural numbers through mod classes}
+\subsection{Fineness on natural numbers through mod classes}\label{example_naturals}
 
@@
 
-In this section we summarize ideas that did not work.
-
-\subsection{Transport of statements}
-
-TODO(GC): Make this an example of how measure zero is not equigranularity
-
-Consider the Borel sigma-algebra on $(0,1)$.
-Let $\stmt_1 = (1/2,1) \setminus \mathbb{Q}$, and let $\stmt = (0,1/3)$.
-The hope is that there would exist some $\stmt_2\narrower\stmt_1$ with $\stmt_2\eqgran\stmt$.
-In Lebesgue measure (and perhaps any reasonable measure of fineness), we would have $\stmt\finer\stmt_1$.
-However, if we try to ``move" $\stmt$ to an equigranular copy inside $\stmt_1$, it will have infinitely many ``holes" - if we wish to use other points of $\stmt_1$ to fill them, there are numerous countable subsets which we consider non-equigranular, and so it is unclear if there is any canonical choice of equigranularity class for such an $\stmt_2$.
-
-Note that even two countable subsets of possibilities need not be equigranular.
-For example, if $\{\stmt[x]_i\}_{i=0}^{\infty}$ is a set of countably many (equigranular) possibilities, and $\stmt = \bigOR\limits_{i=0}^{\infty}\stmt[x]_i$, then $\stmt' = \bigOR\limits_{i=0}^{\infty}\stmt[x]_{2i}$ is strictly finer.
-
-Note: I didn't quite recall what we were thinking with this TODO.
-But if we are wanting to show that two measure zero sets can be non-equigranular, we can go much more generally and not rely on a measure with the above example of countable sets.
-
-\subsection{Transport of sequences}
+In this section we give some counterexamples for ideas that may seem reasonable but in the end do not work.
+
+\subsection{Notes on cardinality}
+
+\textbf{Takeaway: as cardinality ``collapses'' all infinite sets into few categories, it does not tell us much in terms of measure and fineness.} Two sets having the same measure does not imply having the same cardinality and vice-versa.
+Two sets with infinitely many possibilities can be equigranular, one finer than the other or not even comparable.
+
+Consider the reals with the standard measure.
+A set of finitely many points points has measure zero.
+The rationals are countable and have measure zero.
+The Cantor set is uncountable and have measure zero.
+Therefore sets of real numbers with all possible cardinality can have measure zero.
+On the other hand, any finite open interval has uncountably many points and non-zero measure.
+Therefore two sets with the same cardinality can have different measure.
+
+Take example \ref{example_naturals}.
+We have $D_{2,0} \eqgran D_{2,1}$, which both have infinitely many elements.
+We also have $D_{2,0} \supset D_{4,0}$, which are both infinite sets but the first will have to be coarser than the second.
+Now consider the set of prime numbers.
+It is not clear how that can be compared with the mod classes as it is not expressible with finite intersection or union.
+
+\subsection{Limits and levels of granularity}
+
+\textbf{Takeaway: given a sequence of sets with a limit, the level of granularity of the limit is not uniquely determined by the level of granularity of the members of the sequence.}
 
@@
 
-\subsection{Maximal chains}
+\subsection{Saturated chains}
+
+\textbf{Takeaway: chains of narrowness do not go through all granularity levels.}
 
 The idea was to see whether, instead of requiring totality, we could require that we could always find statements at a desired level of description.
-We wanted to say that a maximal chain in narrowness, that is a sequence of statements ordered by narrowness where no other statement could be added, would touch all levels of descriptions, all equivalence classes in equigranularity.
+We wanted to say that a saturated chain in narrowness, that is a sequence of statements ordered by narrowness where no other statement could be added, would touch all levels of descriptions, all equivalence classes in equigranularity.
 This does not work.
 
-The problem is that infinite sets disrupt the relationship between narrowness and fineness, such that a maximal chain in narrowness is not necessarily a maximal chain in fineness.
+The problem is that infinite sets disrupt the relationship between narrowness and fineness, such that a saturated chain in narrowness is not necessarily a saturated chain in fineness.
 For example, if we have a countable set of possibilities (e.g.
-the natural numbers), a maximal chain in narrowness can start from the empty set, start by adding possibilities one be one until we end with the full set.
+the natural numbers), a saturated chain in narrowness can start from the empty set, start by adding possibilities one be one until we end with the full set.
 The set with all possibilities except one, for example, cannot be added to the chain, but still represents a level of description that is in between all the finite sets and the full infinite set.
-The only case where we can construct maximal chains is precisely when the possibilities are finite, and is therefore not interesting.
+The only case where we can construct saturated chains is precisely when the possibilities are finite, and is therefore not interesting.
 
@@
 
-TODO(GC): everything needs to be changed to saturated
-
 \begin{defn}
-A theoretical domain $\tdomain$ is said to have the \textbf{maximal chains property} (MCP) if given any chain $S\subseteq\tdomain$ totally ordered by narrowness ($\narrower$), the following holds.
-There exists a maximum chain $S'\supseteq S$, and a maximal chain $C$ in the set of equigranularity equivalence classes, such that for all statements $s \in S'$, there exists an equivalence class $\overline{c}\in C$ with $s\in \overline{c}$.
+A theoretical domain $\tdomain$ is said to have the \textbf{saturated chains property} (SCP) if given any chain $S\subseteq\tdomain$ totally ordered by narrowness ($\narrower$), the following holds.
+There exists a maximum chain $S'\supseteq S$, and a saturated chain $C$ in the set of equigranularity equivalence classes, such that for all statements $s \in S'$, there exists an equivalence class $\overline{c}\in C$ with $s\in \overline{c}$.
 \end{defn}
@@
 \begin{prop}\label{chainimpliesuniform}
-Let $\tdomain$ be a theoretical domain with the maximal chains property, and suppose all possibilities are comparable. Then they must be equigranular.
+Let $\tdomain$ be a theoretical domain with the saturated chains property, and suppose all possibilities are comparable. Then they must be equigranular.
 \end{prop}
@@
 Consider the two chains $\{\impossibility, p\}$ and $\{\impossibility, q\}$.
-By the MCP, these may be extended to maximal chains in narrowness, with corresponding maximal chains in fineness equivalence classes.
+By the SCP, these may be extended to saturated chains in narrowness, with corresponding saturated chains in fineness equivalence classes.
 Then since possibilities are successors of the impossibility, the fineness equivalence classes represent by $p$ and $q$ must both be an immediate successor of the equivalence class representing the impossibility.
@@
 \begin{prop}
-Let $\tdomain$ be a theoretical domain with the maximal chains property. Then they must have finitely many possibilities.
-\end{prop}
-\begin{proof}
-Let $\tdomain$ be a theoretical domain with the maximal chains property and such that the set $X$ of possibilities is infinite.
+Let $\tdomain$ be a theoretical domain with the saturated chains property. Then they must have finitely many possibilities.
+\end{prop}
+\begin{proof}
+Let $\tdomain$ be a theoretical domain with the saturated chains property and such that the set $X$ of possibilities is infinite.
 Let $X' = \{x_i\}_{i=0}^{\infty} \subseteq X$ be a countable subset.
@@
 
-To show that $C$ is maximal, take a statement $\hat{\stmt} \narrower \stmt$ such that $\hat{\stmt} \notin C$.
+To show that $C$ is saturated, take a statement $\hat{\stmt} \narrower \stmt$ such that $\hat{\stmt} \notin C$.
 Let $A \subseteq X'$ be the set of possibilities compatible with $\hat{\stmt}$.
@@
 We have $\hat{\stmt} \nnarrower \stmt_{j}$ and $\stmt_{i} \nnarrower \hat{\stmt}$ since each is compatible with a possibility incompatible with the other.
-Since we cannot add other statement, $C$ is maximal.
-
-Now we show that, even though $C$ is maximal, it does not go through all levels of granularity.
+Since we cannot add other statement, $C$ is saturated.
+
+Now we show that, even though $C$ is saturated, it does not go through all levels of granularity.
 Let $x \in X'$ and consider $\stmt \AND \NOT x$.
 This is finer than $\stmt$ and is coarser than all other statements statements in $C$ since they are all compatible with finitely many possibilities.
-Therefore MCP does not hold.
-Therefore, if MCP holds, the domain cannot have infinitely many possibilities.
+Therefore SCP does not hold.
+Therefore, if SCP holds, the domain cannot have infinitely many possibilities.
 \end{proof}

### 2021-05-04 Overviews/Informational granularity/InformationGranularity.tex
InfGran: review with Christine up to 2.1 included

@@
 Note that often in physics, geometric structures are recovered from other geometric structures through the idea of emergence.
-For example, substantial effort is put into constructing theories where space-time emerges from other more fundamental structures.
+For example, substantial effort is put into constructing theories where space-time emerges from other more fundamental geometric structures.
 This does not solve the problem: it simply pushes it backwards.
-We are interested in the more fundamental problem of how the first geometric structure, and in fact any geometric structure, is defined.
-
-The idea is that while experimental verifiability keeps track of which statements in the theory are experimentally verifiable, informational granularity keeps track of which provide a finer level of description.
+We are interested in the more fundamental problem of how the first geometric structure, and in fact any geometric structure, arises in physics.
+
+The idea is that while experimental verifiability keeps track of which statements in the theory are experimentally verifiable, information granularity keeps track of which provide a finer level of description.
 \textbf{The working hypothesis is that this, and only this, is what geometric and measure theoretic structures capture: how coarse or fine grained the statements are within a scientific theory.} Finding the fundamental axioms associated to the notion of information granularity is therefore the central problem.
@@
 
-We then suppose we have a way to compare any two statements and decide which one provides a description at a finer level of granularity.
+We then suppose we have a way to compare any two statements and decide whether one provides a description at a finer level of granularity.
 This defines an additional preorder $\finer : \tdomain \times \tdomain \to \mathbb{B}$.

### 2021-05-04 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: more sections added

@@
 \usepackage{assumptionsofphysics}
+\usepackage{tikz}
+\usetikzlibrary{shapes,snakes}
 \usepackage{hyperref}
@@
 
+\newcommand{\marginleft}[1] {\reversemarginpar\marginpar{#1}}
+\newcommand{\marginright}[1] {\normalmarginpar\marginpar{#1}}
+
 \title{Bare minimum: set theory}
@@
 \begin{abstract}
-This note presents a condensed summary of set theory.
-
-Bare minima are not meant to be either didactic or complete.
-They are meant to condense areas of knowledge in a few pages and function as a knowledge check, refresher and/or reference.
-They collect what we feel is the bare minimum of a subject an intellectually curious mind should be aware of, particularly in the context of foundational questions.
+This note presents a condensed summary of set theory which can function as a crash course, refresher and/or reference.
+Bare minima are meant to give a rough overview, by no means complete, of the subject to the intellectually curious, particularly in the context of foundational questions in physics.
 This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
@@
 We will start with naive set theory, which is the best setting for defining all the basic notions and getting a sense of how the framework works.
-Then we will turn our attention to axiomatic set theory to get a sense of what problems it solves and how, and whether those solutions are aligned with the aims of the foundations of physics.
+Then we will turn our attention to axiomatic set theory to get a sense of what problems it solves and how, and whether those solutions mesh with foundational goals  in physics.
 
@@
 
-\begin{defn}[Common sets] The \textbf{empty set} $\emptyset$ is the set with no elements.
+\begin{defn}[Common sets]
+	 The \textbf{empty set}
+	 \marginleft{Common sets: $\emptyset$, $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{R}$, $\mathbb{C}$}
+	  $\emptyset$ is the set with no elements.
 	 The \textbf{set of natural numbers} $\mathbb{N} = \{ 0, 1, 2, ...
@@
 The \textbf{set of complex numbers} $\mathbb{C}$.
+A \textbf{universe} $U$ is the set that, within a specific context, contains all entities under study.
 \end{defn}
@@
 	Let $A$ and $B$ be two sets.
-If all elements in $A$ are also contained in $B$, we say $A$ is a \textbf{subset} of $B$, noted $A \subseteq B$.
+\marginleft{Subset, superset: $\subseteq$, $\supseteq$} If all elements in $A$ are also contained in $B$, we say $A$ is a \textbf{subset} of $B$, noted $A \subseteq B$.
 In this case, we also say $B$ is a \textbf{superset} of $A$, noted $B \supseteq A$.
@@
 \begin{defn}[Power set]
-	Given a set $A$, its \textbf{power set}, noted $\mathcal{P}(A)$ or $2^{A}$, is the set of all subsets of $A$. That is, $\mathcal{P}(A) = \{B \, | \, B \subset A \}$.
+	Given a set $A$, its \textbf{power set}, \marginleft{Power set: $2^{A}$, $\mathcal{P}(A)$} noted $\mathcal{P}(A)$ or $2^{A}$, is the set of all subsets of $A$.
+That is, $\mathcal{P}(A) = \{B \, | \, B \subset A \}$.
 \end{defn}
@@
 \begin{defn}[Set operations]
-	We define the following operations between two sets $A$ and $B$:
+	We define \marginleft{Set operations: $\cup$, $\cap$, $\setminus$, $^{\complement}$} the following operations between two sets $A$ and $B$:
 	\begin{description}
-		\item[Union.] Noted $A \cup B$, the union of $A$ and $B$ is the set of all elements that are contained by $A$, $B$ or both.
-		\item[Intersection.] Noted $A \cap B$, the intersection of $A$ and $B$ is the set of all elements that are contained by both.
+		\item[Union.]Noted $A \cup B$, the union of $A$ and $B$ is the set of all elements that are contained by $A$, $B$ or both.
+		\marginright{
+			\def\setA{ (-1,0) circle (2) }
+			\def\setB{ (1,0) circle (2) }
+			\begin{tikzpicture}[scale = 0.3]
+				\filldraw[black!10] \setA;
+				\filldraw[black!10] \setB;
+				\draw \setA; 
+				\draw \setB;
+				\node at (-4.7,0) {$A\!\cup \! B$};
+			\end{tikzpicture} 
+			\begin{tikzpicture}[scale = 0.3]
+				\begin{scope}
+					\clip \setA;
+					\filldraw[black!10] \setB;
+				\end{scope}
+				\draw \setA; 
+				\draw \setB;
+				\node at (-4.7,0) {$A\!\cap \! B$};
+			\end{tikzpicture} 
+			\begin{tikzpicture}[scale = 0.3]
+				\filldraw[black!10] \setA;
+				\filldraw[white] \setB;
+				\draw \setA; 
+				\draw \setB;
+				\node at (-4.7,0) {$A\!\setminus \! B$};
+			\end{tikzpicture} 
+			\def\setU{ (-3,-2) rectangle (3,2) }
+			\def\setA{ (0,0) circle (1.8) }
+			\begin{tikzpicture}[scale = 0.3] 
+				\filldraw[black!10] \setU;
+				\filldraw[white] \setA;
+				\draw \setU;
+				\draw \setA;
+				\node at (-5.3,0) {$A^{\complement}$};
+			\end{tikzpicture} 
+		}\item[Intersection.] Noted $A \cap B$, the intersection of $A$ and $B$ is the set of all elements that are contained by both.
 		\item[Set difference (or relative complement).] Noted $A \setminus B$, $A$ minus $B$ is the set of all elements in $A$ that are not contained by $B$.
 		\item[Complement (or absolute complement).] Noted $A^{\complement}$, represents all elements that are not contained in $A$.
-Note that his operation is context specific: we need to know that, within a certain context, $U$ is the set of all elements under study; then the complement of $A \subseteq U$ is $A^{\complement} = U \setminus A$.
+Note that this operation is context specific: we need to know that, within a certain context, $U$ is the set of all elements under study; then the complement of $A \subseteq U$ is $A^{\complement} = U \setminus A$.
 	\end{description}
@@
 
+\marginright{~\\
+	\begin{tikzpicture}[scale = 0.3] 
+		\node at (-2.5,-4) {$A$};
+		\node at (3,2.3) {$B$};
+		\draw[rounded corners] (0,0.2) rectangle (6,1.5);
+		\draw[rounded corners] (-0.2,0) rectangle (-1.5,-8);
+		\draw[rounded corners] (0,0) rectangle (6,-8);
+
+		\def\circlepath{circle (0.45)}
+		\def\rectanglepath{ ++(-0.45,-0.45)  -- ++(0,0.9)  -- ++(0.9,0) -- ++(0,-0.9) -- ++(-0.9,0)}
+		\def\diamondpath{ ++(-0.5,0)  -- ++(0.5,0.5)  -- ++(0.5,-0.5) -- ++(-0.5,-0.5) -- ++(-0.5,0.5)}
+
+		% B elements
+		\draw (1,0.85) \circlepath;
+		\draw (3,0.85) \rectanglepath;
+		\draw (5,0.85) \diamondpath;
+
+		\definecolor{a}{rgb}{0.945, 0.415, 0.439}
+		\definecolor{b}{rgb}{0.694, 0.847, 0.466}
+		\definecolor{c}{rgb}{0.549, 0.862, 0.854}
+		\definecolor{d}{rgb}{0.301, 0.301, 0.301}
+
+		% A elements
+		\fill[a] (-0.85, -1) circle (0.5);
+		\fill[b] (-0.85, -3) circle (0.5);
+		\fill[c] (-0.85, -5) circle (0.5);
+		\fill[d] (-0.85, -7) circle (0.5);
+
+		% AxB elements
+		\filldraw[fill=a] (1, -1) \circlepath;
+		\filldraw[fill=b] (1, -3) \circlepath;
+		\filldraw[fill=c] (1, -5) \circlepath;
+		\filldraw[fill=d] (1, -7) \circlepath;
+		\filldraw[fill=a] (3, -1) \rectanglepath;
+		\filldraw[fill=b] (3, -3) \rectanglepath;
+		\filldraw[fill=c] (3, -5) \rectanglepath;
+		\filldraw[fill=d] (3, -7) \rectanglepath;
+		\filldraw[fill=a] (5, -1) \diamondpath;
+		\filldraw[fill=b] (5, -3) \diamondpath;
+		\filldraw[fill=c] (5, -5) \diamondpath;
+		\filldraw[fill=d] (5, -7) \diamondpath;
+
+		\node at (3,-9) {$A \times B$};
+	\end{tikzpicture} 
+}
 \begin{defn}[Cartesian product]
-	Given two sets $A$ and $B$, the \textbf{Cartesian product} $A \times B = \{ (a,b) \, | \, \forall a \in A, b \in B \}$ is the set of all possible ordered pairs between the elements of $A$ and $B$.
+	Given two sets \marginleft{Cartesian product: $\times$}  $A$ and $B$, the \textbf{Cartesian product} $A \times B = \{ (a,b) \, | \, \forall a \in A, b \in B \}$ is the set of all possible ordered pairs between the elements of $A$ and $B$.
 \end{defn}
@@
 
+\begin{remark}
+	Relations provide a common foundation for functions (i.e.
+the relation is the set of pairs $(x, f(x))$), equivalences and order (i.e.
+the relation is the set of pairs $(a, b)$ such that $a=b$ or $a\leq b$ respectively).
+This is an example of highly fruitful generalization.
+\end{remark}
+
 \begin{defn}[Binary relation]
@@
 We say $a$ is \textbf{$R$-related} to $b$, noted $aRb$ if $(a,b) \in R$.
-If $A=B$ the relation is said \textbf{homogeneous}, \textbf{heterogeneous} if not.
+If $A=B$ (i.e.
+domain and codomain coincide) the relation is said \textbf{homogeneous}, \textbf{heterogeneous} if not.
 \end{defn}
@@
 the preimage of the full set $B$).
+\end{defn}
+
+\begin{defn}
+	Given a relationship $R$ between $A$ and $B$, the \textbf{converse} or \textbf{inverse} relationship $R^{-1}$ is the relationship where the order of the pairs is inverted.
+That is, $R^{-1} = \{ (b, a) \, | \, (a, b) \in R\}$.
 \end{defn}
@@
 left-total, defined on the whole domain).
+A function is \textbf{bijective} if it is injective and surjective.
+
+	When the functional form $f(a)$ is known, this can be expressed as $a \mapsto f(a)$ (read ``$a$ maps to $f$ of $a$'').
 \end{defn}
+
+\begin{remark}
+	Note that the domain and the codomain can be any set, including scalar products.
+Therefore $f : \mathbb{R} \times \mathbb{R} \to \mathbb{R}$ for which $(x,y) \mapsto \sqrt{x^2 + y^2}$ gives us the Euclidean norm of a two dimensional vector.
+\end{remark}
+
+\begin{defn}[Identity function]
+	Given a set $A$, the identity function $\Id_A : A \to A$ is the function such that $\Id_A(a) = a$ for all $a \in A$.
+\end{defn}
+
+\marginpar[Left]{Right}
+\begin{prop}[Inverse function]
+	Let $f : A \to B$ be a function.
+The corresponding converse relationship $f^{-1}$ is a function if and only if $f$ is bijective.
+In this case $f^{-1} : B \to A$ is also bijective and is called the \textbf{inverse} of $f$.
+\end{prop}
 
@@
 		\item Composition is associative: $R \circ (T \circ S) = (R \circ T) \circ S$
-		\item Composition of functional/injective/serial/surjective relationship is functional/injective/serial/surjective.
-		\item Specifically, if $f : A \to B$ and $g : B \to C$ are (partial) functions, then $g \circ f$ is a (partial) function.
+		\item Converse of composition: $(R \circ S)^{-1} = S^{-1} \circ R^{-1}$
+		\item Composition of functional/injective/serial/surjective relations is \\ functional/injective/serial/surjective.
+		\item If $f : A \to B$ and $g : B \to C$ are (partial) functions, then $g \circ f$ is a (partial) function.
+		\item Let $f : A \to B$ is a bijective function, then $f \circ f^{-1} = \Id_B$ and $f^{-1} \circ f = \Id_A$.
 	\end{itemize}
@@
 
-\begin{defn}[Families]
-
+\begin{defn}[Sets of functions]
+	The set of all possible functions $f : A \to B$ is noted $B^A$.
 \end{defn}
 
+\begin{remark}
+	The notation is due to the cardinality.
+If $|A| = n$ and $|B| = m$, each function is given by choosing an element of $B$ for each element of $A$, that is $m$ choices $n$ times, that is $m^n$.
+The power set is noted $2^A$ because each subset can be identified by a function $U : A \to \{0 , 1\}$, where $f(a)=0$ corresponds to $a \notin U$ and $f(a)=0$ corresponds to $b \in U$.
+\end{remark}
+
+\subsection{Indexed families and sequences}
+
+\begin{defn}[Families]
+	An \textbf{indexed family}, or simply a \textbf{family}, $\{x_i\}_{i \in I}$ is a collection of elements identified by an \textbf{index set} $I$.
+More formally, an indexed family is a tuple $\langle X, I, x \rangle$ where $X$ and $I$ are sets and $x: I \to X$ is a function such that $i \mapsto x_i$.
+A \textbf{sequence} is an family for which the index set is a continuous interval of natural numbers.
+A \textbf{finite sequence} of $n$ elements is noted $\{x_i\}_{i=1}^{n}$ while an \textbf{infinite sequence} is noted $\{x_i\}_{i=1}^{\infty}$.
+\end{defn}
+
+\begin{defn}[Operations on families of sets]
+	Let $\{A_i\}_{i \in I}$ be a family of sets, the operations of intersection $\bigcap\limits_{i \in I} A_i $ and union $\bigcup\limits_{i \in I} A_i$ can be extended to span all members.
+\end{defn}
+
+\begin{defn}[Set-theoretic limit]
+	Let $\{A_i\}_{i=0}^{\infty}$ be a sequence of sets.
+We define the \textbf{limit infimum} as $\liminf\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1}\bigcap\limits_{j \geq i} A_j$ and the \textbf{limit supremum} as $\limsup\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1}\bigcup\limits_{j \geq i} A_j$.
+If the two are equal to the same set then we say the sequence \textbf{converges}, the \textbf{set-theoretic limit} exists and it is equal to that set.
+\end{defn}
+
+\begin{defn}[Monotone sequence]
+	A \textbf{monotone sequence} is a sequence of sets $\{A_i\}_{i=0}^{\infty}$ where each set is a superset (or subset) of the preceding. More specifically, we distinguish four cases:
+	\begin{description}
+		\item[Increasing] when $A_i \subseteq A_{i+1}$ for all $i>0$
+		\item[Decreasing] when $A_i \supseteq A_{i+1}$ for all $i>0$
+		\item[Strictly increasing] when $A_i \subset A_{i+1}$ for all $i>0$
+		\item[Strictly decreasing] when $A_i \supset A_{i+1}$ for all $i>0$
+	\end{description}
+\end{defn}
+
+\begin{prop}
+	All monotone sequences converge.
+For increasing sequences the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1} A_i$, while for decreasing sequences  the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1} A_i$.
+\end{prop}
+
+\subsection{Homogeneous relations, equivalences and orders}
+
+\subsection{Numeric constructions}
+\begin{remark}
+	The constructions that follow exemplifies the standard way numbers are defines in mathematics and it is useful to gain certain insights.
+In the Assumptions of Physics, however, we prefer to understand numbers as order-theoretic concept (i.e.
+the integer numbers can be thought as a linearly ordered set where each element have an immediate successor and an immediate predecessor).
+\end{remark}
+\begin{defn}
+	Let $A$ be a set.
+We define the \textbf{successor} $A^+ = A \cup \{A\}$.
+A set $S$ is called a \textbf{successor set} if it contains the empty set (i.e.
+$\emptyset \in S$ and if contains the successor of every element (i.e.
+if $X \in S$ then $X^+ \in S$).
+The set of \textbf{natural numbers} $\mathbb{N}$ is the intersection of all successor sets.
+\end{defn}
+\begin{remark}
+	The natural numbers so constructed are known as von Neumann ordinals. They are as follows:
 	\begin{itemize}
-	\item families (and operations on families)
-	\item properties (boolean algebra, lattice)
-	\item Ordered pairs and Cartesian product
-	\item distribution rules of cartesian product
-	\item power set
-	\item cardinality
-	\item Notable sets (empty, universal, natural, integers, rationals, reals, complex)
-\end{itemize}
-
-Functions
-\begin{itemize}
+		\item $0 = \{ \} = \emptyset$
+		\item $1 = \{ 0 \} = \{ \emptyset \}$
+		\item $2 = \{0, 1\} = \{ \emptyset, \{ \emptyset\} \}$
+		\item $3 = \{0, 1, 2\} = \{ \emptyset, \{ \emptyset\}, \{ \emptyset, \{ \emptyset\} \} \}$
+		\item $4 = \{0, 1, 2, 3\} = \{ \emptyset, \{ \emptyset\}, \{ \emptyset, \{ \emptyset\} \}, \{ \emptyset, \{ \emptyset\}, \{ \emptyset, \{ \emptyset\} \} \} \}$
 		\item ...
 	\end{itemize}
+\end{remark}
 
@@
 
-Classes, elements, sets and proper classes.
+Introduce paradoxes. Axiomatic set theory is a family of attempt to solve those paradoxes. There are different variations.
+
+Formalization to remove semantic paradoxes. All approaches of axiomatic set theory do that.
+
+%https://en.wikipedia.org/wiki/Zermelo%E2%80%93Fraenkel_set_theory
+
+%https://en.wikipedia.org/wiki/Von_Neumann–Bernays–Gödel_set_theory
+
+Syntactic paradoxes two main strategies.
+Zermelo–Fraenkel (ZFC) only talk about sets and restrict set construction to subsets only.
+von Neumann–Bernays–Gödel (NBG) talks about classes and defines sets as those classes that can be elements of classes.
+Set builder notation works on sets, not classes.
+So, if putting a class in a class creates a contradiction, it's because the class is not a set, but a proper class.
+
+Ugly elements of set theory: all mathematical objects are sets. Numbers are set.
 

### 2021-05-05 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: relation table added

@@
 
-\subsection{Relations and function}
+\subsection{Relations and functions}
 
@@
 \begin{defn}[Binary relation]
-	Given a set $A$, called \textbf{domain}, and a set $B$, called \textbf{codomain}, a \textbf{binary relation} is a set $R \subseteq A \times B$ of ordered pairs.
+	Given a set $A$, \marginleft{Binary\\ relation} called \textbf{domain}, and a set $B$, called \textbf{codomain}, a \textbf{binary relation} is a set $R \subseteq A \times B$ of ordered pairs.
 We say $a$ is \textbf{$R$-related} to $b$, noted $aRb$ if $(a,b) \in R$.
@@
 
+\marginright{
+\begin{tikzpicture}[scale = 0.3] 
+	\node at (-2.5,-4) {$A$};
+	\node at (3,2.3) {$B$};
+	\draw[rounded corners] (0,0.2) rectangle (6,1.5);
+	\draw[rounded corners] (-0.2,0) rectangle (-1.5,-8);
+	\draw[rounded corners] (0,0) rectangle (6,-8);
+
+	\def\circlepath{circle (0.45)}
+	\def\rectanglepath{ ++(-0.45,-0.45)  -- ++(0,0.9)  -- ++(0.9,0) -- ++(0,-0.9) -- ++(-0.9,0)}
+	\def\diamondpath{ ++(-0.5,0)  -- ++(0.5,0.5)  -- ++(0.5,-0.5) -- ++(-0.5,-0.5) -- ++(-0.5,0.5)}
+
+	% B elements
+	\node at (1,0.85) {\includegraphics[scale=0.25]{cat.png}};
+	\node at (3,0.85) {\includegraphics[scale=0.25]{dog.png}};
+	\node at (5,0.85) {\includegraphics[scale=0.25]{swan.png}};
+
+	\definecolor{a}{rgb}{1.0, 1.0, 1.0}
+	\definecolor{b}{rgb}{1, 0.509, 0}
+	\definecolor{c}{rgb}{0.45, 0.225, 0}
+	\definecolor{d}{rgb}{0, 0, 0}
+
+	% A elements
+	\filldraw[fill=a] (-0.85, -1) circle (0.5);
+	\fill[b] (-0.85, -3) circle (0.5);
+	\fill[c] (-0.85, -5) circle (0.5);
+	\fill[d] (-0.85, -7) circle (0.5);
+
+	% AxB elements
+	\node at (1, -1) {$\checkmark$};
+	\node at (1, -3) {$\checkmark$};
+	\node at (1, -5) {};
+	\node at (1, -7) {$\checkmark$};
+	\node at (3, -1) {$\checkmark$};
+	\node at (3, -3) {};
+	\node at (3, -5) {$\checkmark$};
+	\node at (3, -7) {$\checkmark$};
+	\node at (5, -1) {$\checkmark$};
+	\node at (5, -3) {};
+	\node at (5, -5) {};
+	\node at (5, -7) {$\checkmark$};
+
+	\node at (3,-9) {$R$};
+\end{tikzpicture} 
+}
+
 \begin{defn}[Image and preimage]
-	If $X \subseteq A$, the \textbf{image of $X$ under $R$} is the set of all elements in $B$ that are related to at least one element in $X$.
+	If $X \subseteq A$, \marginleft{Image and\\ preimage} the \textbf{image of $X$ under $R$} is the set of all elements in $B$ that are related to at least one element in $X$.
 The \textbf{image of $R$} is set of all elements in $B$ that are related to at least one element in $A$ (i.e.
@@
 
-\begin{defn}
-	Given a relationship $R$ between $A$ and $B$, the \textbf{converse} or \textbf{inverse} relationship $R^{-1}$ is the relationship where the order of the pairs is inverted.
+\begin{defn}[Converse relationship]
+	Given a relationship  \marginleft{Converse: $R^{-1}$} $R$ between $A$ and $B$, the \textbf{converse} or \textbf{inverse} relationship $R^{-1}$ is the relationship where the order of the pairs is inverted.
 That is, $R^{-1} = \{ (b, a) \, | \, (a, b) \in R\}$.
@@
 
-\marginpar[Left]{Right}
 \begin{prop}[Inverse function]

### 2021-05-06 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: added more material

@@
 \documentclass{article}
+
+\usepackage[left=1.8in,right=1.8in,top=.6in,bottom=1in]{geometry}
 
@@
 \usepackage{tikz}
-\usetikzlibrary{shapes,snakes}
 \usepackage{hyperref}
@@
 	 The \textbf{empty set}
-	 \marginleft{Common sets: $\emptyset$, $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{R}$, $\mathbb{C}$}
+	 \marginleft{Common sets: $\emptyset$, $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$, $\mathbb{C}$}
 	  $\emptyset$ is the set with no elements.
@@
 \subsection{Basic definitions}
+
+\marginright{~\\
+	\def\setA{ (-1,0) circle (2) }
+	\def\setB{ (1,0) circle (2) }
+	\begin{tikzpicture}[scale = 0.3]
+		\draw (0,0) circle (3); 
+		\draw (-0.5,-0.5) circle (1.5);
+		\node at (-4.7,0) {$A \subset B$};
+		\node at (-0.5,-0.5) {$A$};
+		\node at (1.3,1.3) {$B$};
+	\end{tikzpicture} 
+}
 
@@
 
-\begin{defn}[Set operations]
-	We define \marginleft{Set operations: $\cup$, $\cap$, $\setminus$, $^{\complement}$} the following operations between two sets $A$ and $B$:
-	\begin{description}
-		\item[Union.]Noted $A \cup B$, the union of $A$ and $B$ is the set of all elements that are contained by $A$, $B$ or both.
 \marginright{
+	~\\
 	\def\setA{ (-1,0) circle (2) }
 	\def\setB{ (1,0) circle (2) }
+	\vspace{1pt}
 	\begin{tikzpicture}[scale = 0.3]
@@
 	\end{tikzpicture} 
+	\vspace{1pt}
 	\begin{tikzpicture}[scale = 0.3]
@@
 	\end{tikzpicture} 
-		}\item[Intersection.] Noted $A \cap B$, the intersection of $A$ and $B$ is the set of all elements that are contained by both.
+}
+
+\begin{defn}[Set operations]
+	We define \marginleft{Set operations: $\cup$, $\cap$, $\setminus$, $^{\complement}$} the following operations between two sets $A$ and $B$:
+	\begin{description}
+		\item[Union.]Noted $A \cup B$, the union of $A$ and $B$ is the set of all elements that are contained by $A$, $B$ or both.
+		\item[Intersection.] Noted $A \cap B$, the intersection of $A$ and $B$ is the set of all elements that are contained by both.
 		\item[Set difference (or relative complement).] Noted $A \setminus B$, $A$ minus $B$ is the set of all elements in $A$ that are not contained by $B$.
@@
 	\end{description}
+
+	Let $X \subseteq \mathcal{P}(A)$ be a set of subsets of $A$. The operations of union $\bigcup\limits_{Y \in X} Y$ and intersection $\bigcap\limits_{Y \in X} Y$ can be extend to span all elements.
 \end{defn}
@@
 
-\begin{defn}[Binary relation]
-	Given a set $A$, \marginleft{Binary\\ relation} called \textbf{domain}, and a set $B$, called \textbf{codomain}, a \textbf{binary relation} is a set $R \subseteq A \times B$ of ordered pairs.
-We say $a$ is \textbf{$R$-related} to $b$, noted $aRb$ if $(a,b) \in R$.
-If $A=B$ (i.e.
-domain and codomain coincide) the relation is said \textbf{homogeneous}, \textbf{heterogeneous} if not.
-\end{defn}
-
-\marginright{
+\marginright{~\\
 	\begin{tikzpicture}[scale = 0.3] 
@@
 	\end{tikzpicture}
+
+	\begin{tikzpicture}[scale = 0.3] 
+		\node at (-2.5,-3) {$B$};
+		\node at (4,2.3) {$A$};
+		\draw[rounded corners] (0,0.2) rectangle (8,1.5);
+		\draw[rounded corners] (-0.2,0) rectangle (-1.5,-6);
+		\draw[rounded corners] (0,0) rectangle (8,-6);
+
+		\def\circlepath{circle (0.45)}
+		\def\rectanglepath{ ++(-0.45,-0.45)  -- ++(0,0.9)  -- ++(0.9,0) -- ++(0,-0.9) -- ++(-0.9,0)}
+		\def\diamondpath{ ++(-0.5,0)  -- ++(0.5,0.5)  -- ++(0.5,-0.5) -- ++(-0.5,-0.5) -- ++(-0.5,0.5)}
+
+		% B elements
+		\node at (-0.85, -1) {\includegraphics[scale=0.25]{cat.png}};
+		\node at (-0.85, -3) {\includegraphics[scale=0.25]{dog.png}};
+		\node at (-0.85, -5) {\includegraphics[scale=0.25]{swan.png}};
+
+		\definecolor{a}{rgb}{1.0, 1.0, 1.0}
+		\definecolor{b}{rgb}{1, 0.509, 0}
+		\definecolor{c}{rgb}{0.45, 0.225, 0}
+		\definecolor{d}{rgb}{0, 0, 0}
+
+		% A elements
+		\filldraw[fill=a] (1,0.85) circle (0.5);
+		\fill[b] (3,0.85) circle (0.5);
+		\fill[c] (5,0.85) circle (0.5);
+		\fill[d] (7,0.85) circle (0.5);
+
+		% AxB elements
+		\node at (1, -1) {$\checkmark$};
+		\node at (3, -1) {$\checkmark$};
+		\node at (5, -1) {};
+		\node at (7, -1) {$\checkmark$};
+		\node at (1, -3) {$\checkmark$};
+		\node at (3, -3) {};
+		\node at (5, -3) {$\checkmark$};
+		\node at (7, -3) {$\checkmark$};
+		\node at (1, -5) {$\checkmark$};
+		\node at (3, -5) {};
+		\node at (5, -5) {};
+		\node at (7, -5) {$\checkmark$};
+
+		\node at (4,-7) {$R^{-1}$};
+	\end{tikzpicture}
+
 }
+\begin{defn}[Binary relation]
+	Given a set $A$, \marginleft{Binary\\ relation} called \textbf{domain}, and a set $B$, called \textbf{codomain}, a \textbf{binary relation} is a set $R \subseteq A \times B$ of ordered pairs.
+We say $a$ is \textbf{$R$-related} to $b$, noted $aRb$ if $(a,b) \in R$.
+If $A=B$ (i.e.
+domain and codomain coincide) the relation is said \textbf{homogeneous}, \textbf{heterogeneous} if not.
+\end{defn}
 
@@
 \begin{defn}[Relation properties]
-	Let $R$ be a binary relation between $A$ and $B$. We define the following properties:
+	Let $R$ \marginleft{Relation properties} be a binary relation between $A$ and $B$. We define the following properties:
 	\begin{description}
@@
 \begin{defn}[Function]
-	A \textbf{partial function} $f : A \to B$ is a binary relationship between $A$ and $B$ that is functional (i.e.
+	A \textbf{partial function} \marginleft{(Partial) \\ functions: \\ $f : A \to B$} $f : A \to B$ is a binary relationship between $A$ and $B$ that is functional (i.e.
 right-unique).
@@
 \begin{defn}[Identity function]
-	Given a set $A$, the identity function $\Id_A : A \to A$ is the function such that $\Id_A(a) = a$ for all $a \in A$.
+	Given a set $A$, \marginleft{Identity: $\Id_A(a)$} the identity function $\Id_A : A \to A$ is the function such that $\Id_A(a) = a$ for all $a \in A$.
 \end{defn}
@@
 \begin{prop}[Inverse function]
-	Let $f : A \to B$ be a function.
+	Let $f : A \to B$ \marginleft{Inverse: $f^{-1}$} be a function.
 The corresponding converse relationship $f^{-1}$ is a function if and only if $f$ is bijective.
@@
 \begin{defn}[Composition]
-	Let $A$, $B$ and $C$ be three sets.
+	Let $A$, $B$ and $C$ \marginleft{Function and relation composition: $R \circ S$, $f \circ g$} be three sets.
 Let $R$ be a binary relationship between $A$ and $B$ and $S$ be a binary relationship between $B$ and $C$.
@@
 \begin{prop}[Composition properties]
-	Relation composition follows the following properties:
+	Relation composition \marginleft{Composition properties} follows the following properties:
 	\begin{itemize}
@@
 \begin{defn}[Sets of functions]
-	The set of all possible functions $f : A \to B$ is noted $B^A$.
+	The set \marginleft{Sets of functions: $B^A$} of all possible functions $f : A \to B$ is noted $B^A$.
 \end{defn}
@@
 
+TODO: add definition of set closed under operation
+
 \subsection{Indexed families and sequences}
@@
 \begin{defn}[Families]
-	An \textbf{indexed family}, or simply a \textbf{family}, $\{x_i\}_{i \in I}$ is a collection of elements identified by an \textbf{index set} $I$.
+	An \textbf{indexed family}, \marginleft{Families, sequences: $\{x_i\}_{i \in I}$, $\{x_i\}_{i=1}^{\infty}$} or simply a \textbf{family}, $\{x_i\}_{i \in I}$ is a collection of elements identified by an \textbf{index set} $I$.
 More formally, an indexed family is a tuple $\langle X, I, x \rangle$ where $X$ and $I$ are sets and $x: I \to X$ is a function such that $i \mapsto x_i$.
@@
 \begin{defn}[Operations on families of sets]
-	Let $\{A_i\}_{i \in I}$ be a family of sets, the operations of intersection $\bigcap\limits_{i \in I} A_i $ and union $\bigcup\limits_{i \in I} A_i$ can be extended to span all members.
+	Let $\{A_i\}_{i \in I}$ \marginleft{$\bigcap\limits_{i \in I} A_i $, $\bigcup\limits_{i \in I} A_i$} be a family of sets, the operations of intersection $\bigcap\limits_{i \in I} A_i $ and union $\bigcup\limits_{i \in I} A_i$ can be extended to span all members.
 \end{defn}
@@
 \begin{defn}[Set-theoretic limit]
-	Let $\{A_i\}_{i=0}^{\infty}$ be a sequence of sets.
+	Let $\{A_i\}_{i=0}^{\infty}$ \marginleft{Set limit: \\ $\lim\limits_{i \to \infty} A_i = A$} be a sequence of sets.
 We define the \textbf{limit infimum} as $\liminf\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1}\bigcap\limits_{j \geq i} A_j$ and the \textbf{limit supremum} as $\limsup\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1}\bigcup\limits_{j \geq i} A_j$.
-If the two are equal to the same set then we say the sequence \textbf{converges}, the \textbf{set-theoretic limit} exists and it is equal to that set.
+If the two are equal to the same set then we say the sequence \textbf{converges}, the \textbf{set-theoretic limit} $\lim\limits_{i \to \infty} A_i = A$ exists and it is equal to that set.
 \end{defn}
 
+\marginright{~\\
+	\def\setA{ (-1,0) circle (2) }
+	\def\setB{ (1,0) circle (2) }
+	\begin{tikzpicture}[scale = 0.3]
+		\draw (0,0) ellipse (5 and 4); 
+		\draw (0.5,-0.7) ellipse (3.5 and 2.7);
+		\draw (1.05,-1.3) ellipse (2.2 and 1.6);
+		\draw (1.25,-1.45) ellipse (1.6 and 1.05);
+		\draw (1.3,-1.6) ellipse (1.05 and 0.55);
+		\node at (0,-5) {$A_i \supset A_{i+1}$};
+		\node at (-2.6, 2.1) {$A_1$};
+		\node at (-1.05,0.55) {$A_2$};
+		\node at (1.3,-1.6) {$...$};
+	\end{tikzpicture} 
+}
 \begin{defn}[Monotone sequence]
-	A \textbf{monotone sequence} is a sequence of sets $\{A_i\}_{i=0}^{\infty}$ where each set is a superset (or subset) of the preceding. More specifically, we distinguish four cases:
+	A \textbf{monotone sequence} \marginleft{Monotone sequence: \\ $A_1 \subseteq A_2 \subseteq ...$\\ $A_1 \supseteq A_2 \supseteq ...$} is a sequence of sets $\{A_i\}_{i=0}^{\infty}$ where each set is a superset (or subset) of the preceding.
+More specifically, we distinguish four cases:
 	\begin{description}
@@
 
-\begin{prop}
-	All monotone sequences converge.
+\begin{prop}[Monotone convergence]
+	All monotone \marginleft{Monotone convergence} sequences converge.
 For increasing sequences the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1} A_i$, while for decreasing sequences  the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1} A_i$.
@@
 \subsection{Homogeneous relations, equivalences and orders}
+
+\begin{remark}
+	We will cover the notions of orders that are necessary for the definition of ordinals and cardinal numbers. Orders will treated more in detail in a Bare Minimum dedicated to order theory.
+\end{remark}
+
+\begin{defn}[Homogeneous relation properties]
+	Let $R \subseteq A \times A$ \marginleft{Reflexivity, symmetry, transitivity, ...} be a homogeneous binary relation between $A$ and itself. We define the following properties:
+	\begin{description}
+		\item[Reflexive.] Every element $a \in A$ is related to itself. That is, $aRa$ for all $a \in A$.
+		\item[Irreflexive.] No element $a \in A$ is related to itself. That is, $aRa$ for no $a \in A$.
+		\item[Symmetric.] If $a$ is related to $b$, then $b$ is related to $a$. That is, if $aRb$ then $bRa$.
+		\item[Antisymmetric.] If $a$ is related to $b$ and $b$ is related to $a$, then $a$ and $b$ are the same element. That is, if $aRb$ and $aRb$ then $a=b$.
+		\item[Transitive.] If $a$ is related to $b$ and $b$ is related to $c$, then $a$ is related to $c$. That is, if $aRb$ and $bRc$ then $aRc$.
+		\item[Total.] Given two distinct elements, one is related to the other.That is, given $a \neq b$, either $aRb$ or $bRa$.
+	\end{description}
+
+\end{defn}
+
+\begin{defn}[Orders and equivalence]
+	A \textbf{partial order} (noted $\leq$, $\lesssim$, $\preceq$) \marginleft{Order, equivalence: $\leq$, $<$, $\equiv$} is a homogeneous binary relationship that is reflexive, antisymmetric and transitive.
+A \textbf{linear order} or \textbf{total order} is an order that is total.
+An order is \textbf{strict} (noted $<$, $\prec$) if it is irreflexive instead of reflexive.
+
+	An \textbf{equivalence relation} (noted $\equiv$, $\sim$) is a homogeneous binary relationship that is reflexive, symmetric and transitive.
+\end{defn}
+
+\begin{defn}[Partitions]
+	A \textbf{partition} \marginleft{Partitions} of a set $A$ is a collections $P \subset \mathcal{P}(A)$ of disjoint subsets that cover all of $A$. That is:
+	\begin{enumerate}
+		\item $\bigcup\limits_{X \in P} X = A$
+		\item $X \cap Y = \emptyset$ for all $X, Y \in P$ where $X \neq Y$
+	\end{enumerate}
+\end{defn}
+
+\begin{defn}[Equivalence classes and quotient set]
+	Let $A$ \marginleft{Eq.
+classes and quotient set: \\ $[a]_{\sim}, a_{/\sim}, A_{/\sim}$} be a set and $\sim$ an equivalence relation on $A$.
+Let $a \in A$.
+The \textbf{equivalence class of $a$ by $\sim$} is the set $[a]_{\sim} = a_{/\sim} = \{b \in A \, | \, b \sim a \}$ of all the elements of $A$ equivalent to $a$.
+The \textbf{quotient set of $A$ by $\sim$} $A_{/\sim} = \{ a_{/\sim} \, | \, a \in A \}$ all of the equivalences classes by $\sim$.
+\end{defn}
+
+\begin{prop}
+	Every quotient set is a partition, and every partition is a quotient set.
+That is, let $A$ be a set and $\sim$ an equivalence class, then $A_{/\sim}$ is a partition.
+Conversely, let $P$ be a partition of $A$, then $\sim \; = \{ (a,b) \, | \, a, b \in X \text{ for some } X \in P \} $ is an equivalence relation and $A_{/\sim} = P$.
+\end{prop}
+
+%\begin{prop}[Set inclusion as ordering]
+%	Set inclusion \marginleft{$\subseteq$ and $\supseteq$ are partial orders} (either $\subseteq$ or $\supseteq$) is a partial order.
+Strict set inclusion (either $\subset$ or $\supset$) is a strict partial order.
+%\end{prop}
 

### 2021-05-10 Overviews/Process entropy/ProcessEntropy.tex
ProcEntropy: assembling material from other documents

\documentclass[10pt, onecolumn, longbibliography, nofootinbib]{revtex4-2}

\usepackage{assumptionsofphysics}

\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\frenchspacing

% Usual (decimal) numbering
\renewcommand{\thesection}{\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}

% Fix references
\makeatletter
\renewcommand{\p@subsection}{}
\renewcommand{\p@subsubsection}{}
\makeatother

\begin{document}

\title{Assumptions of Physics blueprint: counting evolutions as a foundation for entropy and system/state definition}
\author{Gabriele Carcassi}
\email{carcassi@umich.edu}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\author{Christine Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\date{\today}

\begin{abstract}
    This paper presents ideas and progress towards a more general reconceptualization of entropy that can be used to recover and extend thermodynamics and statistical mechanics. This would address foundational questions such as: why does entropy posses both objective qualities (i.e. it is a measurable property) and in (i.e. it is an ensemble property)? What is entropy maximization and why is so successful in many fields, not just physics? Is there a general notion of equilibrium that includes both dynamical and thermodynamical equilibria in a way that is relativistically invariant? The working hypothesis is that entropy is the logarithm of the count of the possible evolutions of a system within a specific process. That is, given the description of system at the desired granularity level, we count all the way the system can evolve in time that are compatible with that description. This switch from counting states to counting evolutions makes the concept more relativistically sound and allows to properly keep track of how the information about the system is transported back and forward in time by the process, which is essential to apply the concept in a more general setting. The objective is to formulate a precise mathematical framework that makes this intuition precise and recovers the usual entropies in specific settings.

	This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
\end{abstract}

\maketitle

\section{Introduction}

\emph{This paper presents the goal and the current status of our research. It is intended to gather early feedback, including pointers to other relevant work. Feedback is welcome and encouraged.}

The overall goal of Assumptions of Physics is to identify a handful of physical principles from which the basic laws can be rigorously derived. Part of this work is establishing a formal framework for states and processes, of which the definition of entropy is a crucial aspect. We will not assume specific knowledge of the rest of our framework\cite{aop-book,aop-overview-verifiability}, though ultimately the goal is to integrate this work with the rest. As usual, a formal construction from scratch will force us to make explicit the physical requirements that are baked into the use of the common concepts and provide us with a framework that is, hopefully, still valid when those assumptions fail.

Previous work focused on showing how requiring experimental verifiability of a theory leads to topological and $\sigma$-algebraic structures within that theory. We have also seen how the ability to compare the level of granularity within the theory is connected to geometric and measure theoretic structures, giving us a way to quantify the information contained in each statement of the theory. We now want to characterize the relationship between the granularity of the descriptions at different times (i.e their information content. \textbf{The working hypothesis is that entropy captures how coarse or fine grained is the description of a particular system within a particular process. It does so by measuring the logarithm of the count of the possible evolutions that are compatible.} Understanding what are the appropriate basic definitions, how do they map to the standard one and how they are required for the very definitions of systems and states is therefore the central problem of this paper.

To give a sense of how this would work, we start with defining a \textbf{process domain} $\tdomain[P]$. This is a particular instance of our notion of a theoretical domain, and it will represent all meaningful statements, all descriptions at all times and at all scales, that apply to a particular process. Mathematically, it is a $\sigma$-complete Boolean algebra, which means it is closed under negation, countable disjunction (i.e. logical OR) and countable conjunction (i.e. logical AND). It will also come equipped with a partial order $\narrower$ such that $\stmt_1 \narrower \stmt_2$ if $\stmt_1 \OR \stmt_2 \equiv \stmt_2$ that indicates whether one statement gives a \textbf{narrower}, more specific, description than the other. For example:
\begin{itemize}
    \item \statement{The position of the object is between 0 and 1 meters} $\narrower$ \statement{The position of the object is between 0 and 1 kilometers}
    \item \statement{The fair die landed on 1} $\narrower$ \statement{The fair die landed on 1 or 2}
    \item \statement{The first bit is 0 and the second bit is 1} $\narrower$ \statement{The first bit is 0}
\end{itemize}

Additionally, we will have a way to compare any two statements and decide whether one provides a description at a finer level of granularity. Mathematically, we have an additional preorder $\finer : \tdomain \times \tdomain \to \mathbb{B}$. Saying $\stmt_1 \finer \stmt_2$ means that the description provided by $\stmt_1$ is \textbf{finer}, gives more information, is more precise, than the description provided by $\stmt_2$. We say that two descriptions are \textbf{equigranular}, noted $\stmt_1 \eqgran \stmt_2$, if they provide the same level of detail. For example:
\begin{itemize}
    \item \statement{The position of the object is between 0 and 1 meters} $\finer$ \statement{The position of the object is between 2 and 3 kilometers}
    \item \statement{The fair die landed on 1} $\finer$ \statement{The fair die landed on 3 or 4}
    \item \statement{The first bit is 0 and the second bit is 1} $\finer$ \statement{The third bit is 0}
\end{itemize}
In these cases, the first statement may not be contained or overlap with the second. We will assume fineness have all the properties it needs such that if a statement $\stmt[u] \in \tdomain[P]$ is chose, we can construct a measure $\mu_{\stmt[u]}$ that quantifies the granularity of a statement in terms of the unit. For example, if $\stmt[u]=$ \statement{the position of the ball is within 0 and 1 meters} and $\stmt=$ \statement{the position of the ball is within 3.5 and 5 meters}, then $\mu_{\stmt[u]}(\stmt) = 1.5$.

Each process domain will include a set of possibilities $E$: these statements are the narrowest possible statement and provide the full description the whole process at all times and at all level of detail. We call \textbf{evolutions} these statements as they effectively describe the evolution of the all systems within the process. We will assume that all evolutions are equigranular (i.e. $e_1 \eqgran e_2$ for all $e_1, e_2 \in E$). The measure is therefore always uniform over the possibilities and therefore characterizing the granularity level of a description is effectively equivalent to counting the evolutions that are compatible with such description. The process entropy $S(\stmt) = \ln \mu (\stmt)$ is defined to bethe logarithm of that count, and therefore the logarithm of the measure.

With these basic definitions we can transport statements back forth in time and see how the process entropy changes. More specifically, for each instant of time $t \in T$ we can define a respective subdomain $\tdomain_t \subseteq \tdomain[P]$ that contains all the descriptions at that particular time. Given a statement $\stmt_0$ at time $t_0$, we can define the best prediction $\stmt_1$ at time $t_1 > t_0$ and find that its process entropy must be greater (i.e. $S(\stmt_1) \geq S(\stmt_0)$). We can also define the best reconstruction $\stmt_{-1}$ at time $t_{-1} < t_0$ and find that its process entropy also must be greater (i.e. $S(\stmt_{-1}) \geq S(\stmt_0)$). Intuitively, information about a system at a particular time can only get worse (or stay the same) as we transport that description forward or backward in time along the process. This relationship is symmetric with time. Similarly we can study how states of a particular system evolve in time, define determinism and reversibility in the terms of the ability within a process to predict or reconstruct states.

The strength of the approach is that these definitions will apply to processes independently of the discipline (i.e. physics, economics, ecology, ...)  or the kind (i.e. deterministic, irreversible, stochastic, classical/quantum, ...). In particular, it will work in dissipative systems or dynamical systems with equilibria, where the standard notion of entropy decreases as it proceeds toward equilibrium. Additionally it reduces to the standard notions in the appropriate settings. For example, if we assume the system is characterized by a Hamiltonian system over time (i.e. the extended phase space charted by $\{q^i, p_i, t\}$ with Hamiltonian flow), the measure reduces to counting the states at each time. If the flow is not Hamiltonian, however, the phase space measure is not the correct measure at all times (i.e. the same precision over $q$ and $p$ at different times corresponds to a different amount of information). On the other hand, if we assume the system over a small $\Delta t$ fluctuates randomly according to a distribution $\rho$, then the counting the possible evolutions means counting permutations. The process entropy, then, can be expressed as $\int \rho \ln \rho$, and the Shannon entropy is recovered.

TOADD: as the Boltzmann entropy fails to take into account interactions and correlations between molecules, the count of the microstates and the Shannon/Gibbs entropy fail to take into account system/environment correlations. Since these correlations may change in time, the same description of the same system will corresponds to different entropy in time.

\section{Overview}

We now give a more detailed overview of the goal and status. This will also cover the conceptual understanding and physical motivation, not just the mathematical aspects. Later sections will present both what has been established to work and what not to work in full mathematical detail.

The overall goal is to describe a generic system as it evolves in time by defining a minimal set of concepts that such a setup must define, thus bringing to light the tacit assumptions underlying state spaces and evolution laws. We want to find notions that are equally valid and useful within different settings (e.g. thermodynamics, dynamical systems, classical mechanics, ...).

An excellent review of the history of thermodynamics and classical statistical mechanism, and therefore of the different concepts of entropy, has been compiled by Uffink (cite UFFINK). As for terminology, by Shannon entropy we will generally refer to the family of expressions based on the $p \log p$ formula. This will include the Shannon entropy, it's extension to the continuous, the Gibbs entropy and the Von Neumann entropy. By postulate entropy we will mean the one defined by the fundamental postulate of statistical mechanics, which is the logarithm of the count of microstates. We will not be discussing the Boltzmann entropy specifically, as this can be understood as a special case of the Gibbs entropy where all particles are identically and independently distributed.

\subsection{Preamble}

Before we start, we list of a set of problems this approach will need to solve. We also include common points of confusions on entropy, thermodynamics and statistical mechanics.

\textbf{Success of entropy outside of physics}. As entropy maximization is increasingly used outside of physics, we need a coherent explanation as to why that happens. The postulate entropy at this level is not insightful: it is unlikely that the primary explanation for ecological distributions is fluctuations of atoms. Therefore we need a principled account for the entropy maximization on more general grounds.

\textbf{Entropy increases for dynamical equilibria}. For a dynamical system, the presence of equilibria (or more in general attractors) leads to entropy decrease, which is the opposite of what one would expect. Take a dissipative system, like a damped harmonic oscillator: areas in phase space will shrink which leads to a lower count of states and therefore a lower postulate entropy; distributions will become more concentrated which leads a lower Shannon entropy. Therefore, as we approach equilibrium, the entropy decreases which is the opposite of what we would expect. This needs to be reconciled.

\textbf{Ontological vs subjective view of entropy}. On the nature of entropy, there are two positions that we find untenable. One position takes entropy to be an objective ``ontological'' quantity, ontological in the sense that, like charge or position, is an inherent property of the single instance of the system, such as total energy or total momentum. This does not work as the entropy is a property of the ensemble: for the Shannon entropy is a functional of the distribution $\rho$ and for the postulate entropy is the volume of a phase space region. The other position takes entropy to be the subjective information one has about the system, which also does not work. The ensemble is connected to possible fluctuations of the system which are measurable and independent of what we may or may not believe about the system. The idea here is that the notion of equilibrium is process dependent: it depends on both the system and the environment. It depends what happens at the boundary. Even isolation is a type of boundary, which is also technically impossible (e.g. we cannot isolate from gravitational interactions). Therefore entropy should be epistemological but objective: it quantifies the level of information (epistemological) that is accessible through a particular process allows (objective).

\textbf{Information content is time dependent}. Both the postulate and the Shannon entropy are time independent. This presents a problem because the same statement at different times can provide different information. Take \statement{Mark is in bed}. If it is 2 am, the statement is very likely to be true, therefore provides little information. If it is 1 pm, the statement is very unlikely to be true, and therefore provides more information. In the same vein, if we have a dissipative system, \statement{the energy of the system is less than 3 J} provides less and less information as time goes on. It should be clear that a proper definition of entropy that would work far from equilibrium will need to take this into account.

\textbf{Equilibria and relativity}. Another problem related to time dependence is that, in general theromdynamic equilibria are not relativistic object. The notion of ``equilibria as nothing is changing'' is not invariant under boost. A volume of gas would be at equilibria only if at rest. For a boosted frame, something would be changing (i.e. the position). This also suggests an automatic interaction with gravitational forces: a volume of gas at rest would follow geodesics for other observers. Another failure of thermodynamic equilibria in a relativistic context is the following. Suppose a volume of gas is at rest for an observer reaches equilibrium at time $t_0$. This means there is a region of space $\Delta x$ that after $t_0$ can be considered at equilibrium. But equal time for a boosted frame is different, so different observers will observe different parts reaching equilibrium at different times.

\textbf{Thermodynamic reversibility vs dynamical reversibility}. There are different notion of reversibility. Thermodynamic reversibility is the existence of a process that ``undoes the change''. This is problematic as this process technically does not exist in many cases. Dynamical reversibility is the ability to reconstruct the initial state. This is also called by some retrodictability or reverse determinism. A dissipative system should be regarded as thermodynamically irreversible, yet is technically retrodictable.

\textbf{Units for entropy}. Units are often stripped out and not discussed, which leave the mathematics with imprecise physical meaning. If entropy is associated to phase space volume, we should note that the unit of phase space volume depends on the number of degrees of freedom of the system. This means that the entropy of two systems with different number of particles is not numerically comparable as the areas are not commensurable. We need a precise account for this in the math.

\textbf{Entropy of a single microstate}. The entropy of a single microstate is minus infinity, and not zero. A single microstate occupies zero entropy and the logarithm of zero is minus infinity. In terms of Gibbs/Shannon entropy, a single microstate would correspond to a delta function. The infinite density at that point contributes minus infinite entropy, and since the distribution is zero everywhere else, there are no other contributions. In quantum mechanics, all pure states have zero von Neumann/Shannon entropy. In a way, that is the problem that quantum mechanics fixes.

\textbf{Entropy over continuous variables}. The definitions of entropy are in general not invariant under coordinate transformations (i.e. change of continuous variables). It is the structure of phase space in classical mechanics and of the quantum Hilbert space that make entropy invariant. It can be shown that classical phase space (i.e. symplectic manifolds) is exactly the structure needed for a coordinate invariant (see \cite{aop-phil-Hamiltonianinformation}).

\subsection{Descriptions and logical relationships}

The first step is to give the most basic mathematical structure we must have to be able to describe a process at all levels of granularity. This consists of a purely logical structure that imposes logical relationships between all possible descriptions.

\textbf{Process domain}. Our starting point is the \textbf{process domain} $\tdomain[P]$, which is the set that contains all possible descriptions of the system at all levels of detail at all possible times. Depending on the process, it will contain statements like \statement{the average volume of the system between 1 and 2 seconds is between 3 and 4 liters} or \statement{the trajectory of the particle is $y=10 \, m - t^2 \cdot 9.80665 \, m/s^2 $}. It does not matter at this point what the statements themselves are, except they must be, at least in line of principle, defined from experimentally well-defined starting points. The translation of these physical requirements into a formal structure has already been carried out in previous work (see \cite{aop-book}). Mathematically, $\tdomain[P]$ is a $\sigma$-complete Boolean algebra\footnote{A set of statements that is closed under negation (NOT), countable disjunction (OR) and countable conjunction (AND).} and it is the closure of the set of experimentally verifiable statements $\edomain[P] \subseteq \tdomain[P]$.\footnote{The set $\edomain[P]$ will form a Heyting algebra which, to be experimentally reachable, must allow a countable basis. The closure $\tdomain[P]$ is in terms of negation and countable disjunction.}

\textbf{Logical relationships}. As $\tdomain[P]$ contains different levels of description, some statements will be more or less specific. For example, we say that \statement{the horizontal position is between 2.5 and 3 meters} is \textbf{narrower} than \statement{the horizontal position is between 2 and 3.5 meters}. Note that whenever the first statement is true, the second one must be true as well. Given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$, we formally capture this relationship by noting $\stmt_1 \narrower \stmt_2$ (read ``is narrower than''). Narrowness can describe relationships at different scales, on different quantities (e.g. \statement{the horizontal position is between 2.5 and 3 meters and the vertical position is between 1 and 1.5 meters} $\narrower$ \statement{the horizontal position is between 2.5 and 3 meters}), constraints between variables at the same time (e.g. \statement{the temperature of the water in the glass is 3.98 $^\circ C$} $\narrower$ \statement{the density of the water in the glass is 1 $g/cm^3$}) or at different times (e.g. \statement{at time 0 s the position is 1 m and the velocity is 1 m/s} $\narrower$ \statement{at time 1 s the position is 2 m}). It is a general tool to capture relationships within the theory which intuitively can also be thought of as implication.\footnote{Technically, implication in classical logic is a truth function as classical logic does not incorporate the idea of different ``possible cases''. What we have is more similar in spirit to semantic consequence in modal logic, though modal logic brings in a lot of undesirable elements we do not want.} Mathematically, a Boolean algebra is also a lattice (i.e. a partially ordered set with supremums and infimums) and $\narrower$ is the ordering relationship for $\tdomain[P]$.

We can define other two logical relationships, which will be useful later. For example, we want to say that \statement{the horizontal position is between 2.5 and 3.5 meters} is \textbf{compatible} with \statement{the horizontal position is between 2 and 3 meters}, meaning that they can be true at the same time. Given two statements $\stmt_1, \stmt_2 \in \tdomain[P]$ we denote $\stmt_1 \comp \stmt_2$ (read ``is compatible with'') if it is possible for them to be both true. Finally, we want to capture whether the truth of one or more statements influences the truth of others. For example, we want to say that \statement{the temperature of the water in the glass is 3.98 $^\circ C$} is \textbf{independent} from \statement{the volume of the water in the glass is 10 $cL$} but is not independent from \statement{the density of the water in the glass is 1 $g/cm^3$}. We note $\stmt_1 \indep \stmt_2$ (read ``is independent of'') when two statements are independent.

\textbf{Possible evolutions, topology and $\sigma$-algebra}. Within our process domain $\tdomain[P]$ we can define the set $E$ of the narrowest possible statements. Its elements give a complete description of our system at all times and therefore we call them \textbf{evolutions} while we call $E \subset \tdomain[P]$ the set of all possible evolutions. In the Assumptions of Physics framework, $E$ is the set of possibilities.

Every statement $\stmt \in \tdomain[P]$ can now be characterized by the set $A(\stmt) \subseteq E$ of all evolutions compatible with $\stmt$, that is all the evolutions for which $\stmt$ will be true. Mathematically,  $A(\edomain[P])$ is a \textbf{topology} over $E$ while $A(\tdomain[P])$ forms a $\sigma$-algebra over $E$, the \textbf{Borel algebra} of $A(\edomain[P])$. Narrowness and compatibility become set relationships in the $\sigma$-algebra: $\stmt_1 \narrower \stmt_2$ if and only if $A(\stmt_1) \subseteq A(\stmt_2)$ (i.e.~all evolutions for which $\stmt_1$ is true are such that $\stmt_2$ is also true) and $\stmt_1 \comp \stmt_2$ if and only if $A(\stmt_1) \cap A(\stmt_2) \neq \emptyset$ (i.e.~two descriptions are compatible if there is an evolution that is compatible with both). This connection gives a direct physical meaning to these foundational set-based mathematical structures.

\begin{table}[h!]
	\centering
	\begin{tabular}{|c|c|p{6cm}|}
		\hline 
		Symbol & Name & Meaning \\ 
		\hline 
		$\tdomain[P]$ & Process domain & The set of all possible descriptions at all times and at all levels of granularity \\ 
		\hline 
		$E \subset \tdomain[P]$ & Possible evolutions & The set of descriptions that give a full account of the process; they correspond to the narrowest statements within $\tdomain[P]$ \\ 
		\hline 
		$A(\stmt)$ & Compatible evolutions & The set of evolutions that are compatible with the statement $\stmt$ \\ 
		\hline 
		$\narrower$ & narrower & A statement is narrower than another $\stmt_1 \narrower \stmt_2$ if the second one is true whenever the first one is \\ 
		\hline 
		$\comp$ & compatible & A statement is compatible with another $\stmt_1 \comp \stmt_2$ if they can be true at the same time \\ 
		\hline 
		$\indep$ & independent & Two statements are independent  $\stmt_1 \indep \stmt_2$ if the truth of one does not influence the truth of the other \\ 
		\hline 
	\end{tabular}
	\caption{Process descriptions and their logical relationships}
	\label{table:logic}
\end{table}

\subsection{Granularity and quantifying evolutions}

Now that we have a way to track all possible descriptions about a process and their logical relationship, we need to characterize their granularity. It turns out that this is equivalent to counting how many evolutions are compatible with each description: the more fine-grained the description, the fewer evolution (i.e. the fewer cases) it will identify.

\textbf{Insufficiency of a single measure}. The next step is to quantify the granularity of the description provided by each statement. Since statements at a finer level of description will put a greater constraint on the set of compatible evolutions, quantifying the granularity of the description of $\stmt$ means quantifying the size of $A(\stmt)$. The statement \statement{at time 0 s the particle is between 0 m and 1 m} is doubly more specific than \statement{at time 0 s the particle is between 0 m and 2 m} precisely because it corresponds to half the evolutions of the second statement.

The naive approach would be to assign a measure over $A(\tdomain[P])$: since this is a $\sigma$-algebra over $E$ and a measure is exactly what one uses to give a size to each set, it would seem we are done. Unfortunately it is not that simple: a single measure can only compare objects of the same dimensionality. If we imagine all evolutions to be points on an $n$-dimensional manifold, if the measure gives finite values for $k$-dimensional regions it will necessarily give measure zero for all regions with lesser dimensionality and infinite measure for those with greater dimensionality. Coarse grained and subsystem descriptions will have different dimensionality as they describe different numbers of degrees of freedom, therefore we need a way to quantify objects at all levels.\footnote{Typically one uses geometrical structures for his purpose. In phase space, the symplectic form allows us to define areas of two-dimensional surfaces, and use those to define volumes or even-dimensional areas. Note that it does not quantify the size of the points, which are implicitly assumed to be of equal size. This is a special case and cannot be the starting point.}

\textbf{Fineness and equigranularity}. Our starting point will be the ability to compare two statements and decide which one gives a finer description of the process. Formally, we have preorder $\finer$ on $\tdomain[P]$ which, given two statements, tells us whether the description of one is \textbf{finer}, more refined, than the other. Two statements are \textbf{equigranualar}, noted $\stmt_1 \eqgran \stmt_2$, if they provide the same level of description. For example, we can say that \statement{the trajectory of the particle is $y=t \cdot 1 \, m/s$} $\eqgran$ \statement{the trajectory of the particle is $y=t \cdot 2 \, m/s$} and that \statement{the position at time 0 sec is between 0 and 1 m} $\eqgran$ \statement{the position at time 0 sec is between 1 and 2 m}, even though the second pair is infinitely less discerning than the first. Note that if one statement is narrower than the other (i.e. $\stmt_1 \narrower \stmt_2$) then it is also finer than the other (i.e. $\stmt_1 \finer \stmt_2$) while the converse is not necessarily true.

We say two statements are \textbf{comparable} if one is finer than the other. The preorder $\finer$ is partial because not all statements are comparable to each other. For example, consider the pressure/volume state space for an ideal gas. Comparing \statement{the pressure is 1 kPa and the volume is between 1 and 2 liters} and \statement{the pressure is between 1 and 2 kPa and the volume is 1 liter} would mean saying whether one liter is bigger or smaller than one kPa.\footnote{In phase space, for example, we cannot in general compare $q^i$ and $p_i$, yet we can always compare areas of phase space} The idea of physical dimension would therefore be something that is formally captured mathematically.

The idea is that geometrical, measure theoretic, information theoretic and probabilistic structures all descend from this more fundamental structure, the partial order that describes this ``information granularity''. We note that all these structures can be used to determine a size and therefore a partial order on the set. It would be possible, given sufficient conditions, to recover those structures from the order itself. The mathematical detail of how exactly this would work (i.e. what conditions are necessary and sufficient) is out of the scope of this work (see CITE other blueprint). Here we will simply assume that the process domain is such that the geometrical structures can be rederived.

\textbf{Measures of granularity}. To recover measures, we pick a statement $\stmt[u] \in \tdomain[P]$, called unit, and construct a \textbf{measure} $\mu_{\stmt[u]}$ that is normalized (i.e. $\mu_{\stmt[u]}(\stmt[u]) = 1$), monotonic (i.e. $\stmt_1 \finer \stmt_2$ will mean $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$) and additive (i.e. $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$ if $\stmt_1 \ncomp \stmt_2$).\footnote{Differentiating between finite or countable additivity is beyond the scope.} That is, we can quantify how much more or less precise a statement is compared to a fixed statement taken as a unit. If two statements are such that the measure of one with respect to the other is finite and non-zero, we say they are \textbf{finitely comparable}. We can show that if $\stmt[u]_1$ and $\stmt[u]_2$ are finitely comparable we have $ \mu_{\stmt[u]_1}(\stmt) = \mu_{\stmt[u]_1}(\stmt[u]_2) \mu_{\stmt[u]_2}(\stmt)$. Furthermore, since all evolutions give a complete description of the whole process, we will assume they are all equigranular. Therefore if we pick $\stmt[u] \in E$, then $\mu_{\stmt[u]}(\stmt)=\#(A(\stmt))$ would be the counting measure of evolutions compatible with $\stmt$.\footnote{We still need to understand the necessary and sufficient conditions under which such measures would exist and be unique. For the present work, we assume it can be done.}

\textbf{Independence}. If we fix the evolution of one subsystem, it should be intuitive that the possible evolutions of another subsystem depends on the correlation between the two. The maximum number of joint evolutions corresponds to the case were the systems are independent: the choice of evolution for one does not contrain the choice of evolution for the other. In this case, we expect the measure to factorize. Mathematically, we take two subdomains $\tdomain[P]_1 \subset \tdomain[P]$ and $\tdomain[P]_2 \subset \tdomain[P]$. By subdomains we mean a set of descriptions that also form a $\sigma$-complete Boolean algebra. We say the two subdomains are \textbf{independent} if knowledge of one does not constrain knowledge of the other. Mathematically, any pair of statements $(\stmt_1, \stmt_2) \in \tdomain[P]_1 \times \tdomain[P]_2$ from the two domains is independent $\stmt_1 \indep \stmt_2$. As we expect the count of evolution to factorize, we have $ \mu_{\stmt[u]_1 \AND \stmt[u]_2}(\stmt_1 \AND \stmt_2) = \mu_{\stmt[u]_1}(\stmt_1) \mu_{\stmt[u]_2}(\stmt_2)$. In terms of the math, this can be imposed but it would be nice to show that this has to be the case. It can be shown for finite sets. It is not clear whether it must be for infinite ones. Also, note the units of the measures are different. As an analogy to classical phase-space, the area of each degree of freedom can be expressed in units of $\hbar$. The volume of two degrees of freedom, however, is in units of $\hbar^2$. This is an additional detail that needs to be fully understood mathematically in a way that is typically glossed over in statistical mechanics.

\textbf{Process entropy}. We are now ready to define the \textbf{process entropy} with respect to $\stmt[u]$ as the quantity $S_{\stmt[u]}(\stmt) = \log \mu_{\stmt[u]}(\stmt)$. This can be thought of as the number of bits, yes/no questions, that separate the level of description of $\stmt[u]$ from the one provided by $\stmt$. Note that process entropy combines linearly for independent systems: $S_{\stmt[u]_1 \AND \stmt[u]_2}(\stmt_1 \AND \stmt_2) = \log \mu_{\stmt[u]_1 \AND \stmt[u]_2}(\stmt_1 \AND \stmt_2) = \log (\mu_{\stmt[u]_1}(\stmt_1) \mu_{\stmt[u]_2}(\stmt_2)) = \log \mu_{\stmt[u]_1}(\stmt_1) + \log \mu_{\stmt[u]_2}(\stmt_2) = S_{\stmt[u]_1}(\stmt_1) + S_{\stmt[u]_2}(\stmt_2)$. Also note that there is a unit dependence, so one can't simply compare the entropy of two different systems as they would be, most likely, expressed in different units.  It typically does not pose a problem because we compare entropy of different description of the same object. This issue already exists with the other entropies, we are simply making in explicit so that we can study it.\footnote{The phase space volume will be in units $\hbar^n$ where $n$ is the number of degrees of freedom, typically $n=3m$ where $m$ are then number of molecules. While $3 < 4$, there is no sense in which $3 \hbar^6$ is smaller than $4 \hbar^{12}$ therefore comparing the entropy of gasses with different number of particles is not well defined.}

\textbf{Recovering the postulate entropy}. Entropy as the logarithm of the state count can be easily recovered as a special case. Suppose the process is deterministic and reversible in the sense that the state at one time identifies the whole evolution. Suppose that all state descriptions at one time are equigranular. In that case, the count of the evolution corresponds to the count of states and therefore the process entropy coincides with the postulate entropy.

\textbf{Recovering the Shannon entropy}. One way to recover the Shannon entropy (though probably not the only one) is to assume we have a process that is fluctuating according to a stable distribution $\rho(s)$. That is, our time resolution is fixed at a scale $\Delta t$, the state $s$ keeps changing within that interval. The distribution $\rho(s)$ represents how often the system will visit the state $s$ within that timeframe. To calculate the process entropy we need to calculate how many possible evolutions we would have within $\Delta t$ that satisfy $\rho(s)$. An evolution is a sequence of states, one for each instant in time. If we assume the evolution is continuous, countably many states will suffice to identify one evolution. Therefore we are looking to count the possible permutations of a countably infinite sequence with recurrence set by $\rho(s)$. But this is exactly one way to recover the Shannon entropy. Therefore, in this case, the process entropy coincides with the Shannon entropy.

\textbf{Probability}. Quantifying evolutions allows us to define probability. The idea is a process can be realized differently by each possible evolution. The more evolutions a statement is compatible with, the more it is likely to happen. For example, consider a process that implements a coin flip. If the coin (and the process flipping the coin) is fair, we must have that half the evolutions are compatible with \statement{the coin is head} while the other half are compatible with \statement{the coin is tails}.  Therefore we define that the \textbf{probability} of $\stmt_2$ given $\stmt_1$ is
\begin{equation}
	P(\stmt_2 | \stmt_1) = \mu_{\stmt_1}(\stmt_1 \AND \stmt_2) = \frac{\mu_{\stmt[u]}(\stmt_1 \AND \stmt_2)}{\mu_{\stmt[u]}(\stmt_1)}
\end{equation} where $\stmt_1 \AND \stmt_2$ is the logical conjunction (AND) between the two statements and $\stmt[u]$ is finitely comparable to $\stmt_1$. In other words, the probability quantifies the ratio of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$. This can be shown to satisfy the standard axioms of probability.

\begin{table}[h!]
	\centering
	\begin{tabular}{|c|c|p{6cm}|}
		\hline 
		Symbol & Name & Meaning \\ 
		\hline 
		$\finer$ & finer & A statement is finer than another $\stmt_1 \finer \stmt_2$ if it provides a more refined description than the second \\ 
		\hline 
		$\eqgran$ & equigranular & A statement is equigranular to another $\stmt_1 \eqgran \stmt_2$ if they provide the same level of detail \\ 
		\hline 
		$\mu_{\stmt[u]}(\stmt)$ & Measure & Quantifies the precision of a statement $\stmt$ in terms of a reference unit statement $\stmt[u]$; it quantifies how many evolutions are compatible with $\stmt$ compared to the evolutions compatible with $\stmt[u]$  \\ 
		\hline 
		$S_{\stmt[u]}(\stmt)$ & Evolution entropy & Defined as $\log \mu_{\stmt[u]}(\stmt)$; represents the number of bits that separate the level of description of $\stmt$ from the one provided by the unit statement $\stmt[u]$; it quantifies the number of questions needed to go from the level of description of $\stmt$ to the level of description of $\stmt[u]$  \\ 
		\hline 
		$P(\stmt_2 | \stmt_1)$ & Probability & Defined as $\mu_{\stmt_1}(\stmt_1 \AND \stmt_2)$; quantifies the fraction of evolutions compatible with $\stmt_1$ that are also compatible with $\stmt_2$  \\ 
		\hline 
		& Comparable & Two statements are comparable if one is finer than the other \\ 
		\hline 
		& Finitely comparable & Two statements are finitely comparable if the measure of one in terms of the other is finite and non-zero \\ 
		\hline 
	\end{tabular} 
	\caption{Granularity, measures and probability}
	\label{table:states}
\end{table}

\subsection{Time, predictions and reconstructions}

Now that we have a way to keep the logical structure and granularity of all descriptions of a process, we want to understand how the information is transported back and forward in time by the process. That is, once a particular description is verified at a particular time, what can we say about past and future times.

\textbf{Time domains}. As a process extends over time and potentially describes multiple systems, we want to organize the statements into subdomains, one for each moment in time relative to a particular system. Given a \textbf{time parameter} $t \in T \subseteq \mathbb{R}$, we can imagine carving out a \textbf{system domain} $\tdomain_t$ that would correspond to all the statements about the system one can make at the given time. In particular, we define \textbf{one-step process} as one that only has an \textbf{initial domain} $\tdomain_{t_0}$ and a \textbf{final domain} $\tdomain_{t_1}$. Note that $\tdomain_t$ is a $\sigma$-complete Boolean sub-algebra of $\tdomain[P]$ since any logical operation on statements at fixed $t$ is also another statement at fixed $t$.

\textbf{Temporal and scale coarse graining}. Technically, the statements are not defined at an infinitesimal moment $t$, but rather within a finite interval $[t, t + \Delta t]$. This represents the time-scale at which the process will be described, meaning the description within $\tdomain_t$ should not be sensitive to what happens at faster scales. The quantities used to represent states at time $t$, then, should really be thought of as averages within $\Delta t$. In the same way, $\tdomain_t$ will in general represent the system at the chosen level of description, not the narrowest level possible. Statements about the environment or at a finer level of description (i.e. the positions of all molecules for a gas) will not be included.

\textbf{Optimal coarse graining}. In principle, this coarse graining is arbitrary and subjective in the sense that is completely at the mercy of the decision of the scientist. However, not all choices are equivalent. Ideally, we want the process at our level of description to not depend on the faster or smaller scale dynamics. If the dynamics at faster and smaller scale decouples from the one at slower and bigger scale, we can study the latter without worrying about the former. So, for example, we can study the trajectory of a cannonball without worrying about the motion of each individual molecule; we may not be able to predict the exact energy at a future time, as it fluctuates through interactions with the environment, but we may be able to predict it's average over a small $\Delta t$. We should stress here that \textbf{the process itself tells us whether and how the faster/smaller scale decouples decouple from the slower/bigger scale}. Therefore the choice of an \textbf{optimal} coarse graining is not arbitrary or subjective at all: it is determined by the process. The processes at our disposal guide us to carve domain at both temporal and size scale, and finding the optimal level of scale is part of the job of a physicist. At a theoretical level, we need to understand how a process guides us to that choice.

\textbf{Predictions and reconstructions}. A crucial step, then, is to characterize how the information at one time is carried to other times by the process. Suppose we found that the statement $\stmt_0 \in \tdomain_{t_0}$ at time $t_0$ was true. What can we predict of the system at a future time $t \geq t_0$? Any future statement $\stmt \in \tdomain_{t}$ such that $\stmt_0 \narrower \stmt$ will be true if $\stmt_0$ is true. Therefore the best prediction will be the conjunction (the logical AND) of all such statement. In terms of order theory, it will be the narrowest broader statement. As all statements can be thought as sets of evolutions, it will be statement $\stmt \in \tdomain_{t}$ such that $A(\stmt)$ is the smallest set that contains $A(\stmt_0)$.

In a similar way, we can ask what can be reconstructed about the system at a previous time $t \leq t_0$. Any past statement $\stmt \in \tdomain_{t}$ such that $\stmt_0 \narrower \stmt$ will be true if $\stmt_0$ is true. Therefore the best reconstruction will be the conjunction of all such statements, the narrowest broader statement, the smallest set $A(\stmt)$ that contains $A(\stmt_0)$.

Given that the formal definition for predictions and reconstructions are similar, we note $P_{t}(\stmt_0)$ the prediction or recostribution for statement $\stmt_0$ at time $t$.

\textbf{Predictions and reconstructions always have higher entropy}. Prediction and reconstruction are necessarily broader than the statement we start from and therefore they will necessarily be coarser: they cannot give us more information then what we started with. They may either give us the same information or less information. Mathematically, $\stmt_0 \finer P_t(\stmt_0)$ for any $\stmt_0$, $t_0$ and $t$. In this respect, there is no time asymmetry: information can only be diminished as we map it to past or future time.

\textbf{The process dictates the relationship between the information at different moments in time}. How much the information diminishes depends exclusively on the process. Suppose we take a deck of cards and memorize the order and leave it on the table. If sits undisturbed, full knowledge is maintained. If a person takes the top card and puts it at the bottom, full knowledge is still maintained. If a person shuffles the deck, the information about the order of the deck is lost. If a person cuts the deck, the information about the first card is lost, yet the information about the order is not.

\textbf{Natural graining}. As each process defines how information is transported over time, it will define optimal granularity levels for its description. We say $\stmt_0 \in \tdomain_{t_0}$ \textbf{cuts the process along the grain}, or simply \textbf{is a natural cut}, if its prediction and reconstructions do not lose information. That is, $\stmt_0 \eqgran P_t(\stmt_0)$ for all $t$. The certainty and the impossibility are both natural cuts of all process. More work needs to be done to understand the role of natural graining in a process, but the general sense is this would be a key element for a better definition of states and equilibria because of the following properties.

\textbf{Natural graining is automatically relativistic}. Consider a free particle, whose state at a given time $x_0$ is fully determined by position and momentum. This will give us predictions and reconstructions at the same level of granularity for all times, therefore $x_0$ cuts along the grain. It doesn't matter whether the particle is at rest or in motion. It doesn't matter whether ``all properties'' remain the same.

\textbf{Natural graining is compatible with both dynamic and thermodynamic equilibrium}. Continuing with the card deck example, suppose a person takes the deck of card, cuts in half, shuffles the bottom half and orders the top part by suit and rank. This process creates an equilibrium that is, in a way, a mix of dynamic (the top part) and thermodynamic (the bottom part) equilibria. If we apply the process another time, in fact, the top part remains exactly in the same configuration; the bottom part, while it will most likely change, will retain its statistical description.  Each equilibrium is fully identified by the cards in the top part. If we take the full description of the deck at the beginning, the prediction is coarser, therefore it is not a natural cut. The full description of the top half of the deck will also give a coarser prediction: the final order of the top half could have been reached by any initial order of the same cards in the top half. The description of what cards are in the top half, however, gives us exactly the information at equilibrium. Therefore it is a natural cut. This shows that the idea of natural cuts captures both types of behavior, even when they happen at the same time.

\textbf{Natural graining is time independent}. One of the problems with the notion of equilibria is defining when the equilibria is reached. Natural graining is a time independent concept, so we do not need to define when it is reached. If we consider a dynamical equilibrium, we have a basin of attraction for an equilibrium. All evolutions that start in the basin of attraction will tend closer and closer to the equilibrium, without ever reaching it. The natural graining consists in the set of all evolutions taken together. In many cases, we are not interested in a specific evolution. For example, we have a cutoff at a certain time and we are only interested to note which equilibria the system was approaching. Mathematically, what needs to happen is that at the cutoff the description is a natural cut for both the previous and following process. At that point the behavior in the limit is not important, which makes the framework more compatible with realistic scenarios.

\textbf{Natural graining decouples the internal dynamics}. Consider a macroscopic object, like a balloon. Though we know it is formed by on the order of $10^23$ molecules, we can still characterize its state with a few variables. Each molecule itself, maybe formed by a large number of fundamental constituents. At each level, the description is fairly independent of the description of the constituents. Natural graining picks out those descriptions that are effectively independent of the faster/smaller scale.

\subsection{Snapshots, determinism and reversibility}

In the previous section we explored how the information at a particular time moves across a process. Now we want to concentrate on the best possible descriptions of the system at each time. That is, how the descriptions with maximal information at each time relate to each other. This will give us straight forward definitions for determinism and reversibility.

\textbf{Snapshots and trajectories}. For each time domain $\tdomain_t$ we can find the set of possibilities $X_t$, which represent the statements that will give the most precise description at that time; we call these the possible \textbf{snapshots} of our system at that time. A \textbf{trajectory} is a sequence $x(t)$ of snapshots at each time. A particular evolution will identify a trajectory: that is, given $\stmt[e] \in E$ we can find $x(t)$ such that $\stmt[e] \narrower x(t_0) \AND x(t_1) \AND x(t_2) \AND ... $. The converse is not necessarily true because $\tdomain_t$ will typically describe a particular system at a particular level of granularity and a particular time-scale. Therefore a single trajectory may correspond to multiple evolutions, since the evolutions potentially describe the same system at greater accuracy and/or other systems.

\textbf{Snapshots are not states}. One key insight is that the set of possible snapshots $X_t$ may not be the same at all times and may not correspond to the state of the system. First of all $X_t$ will only include those configurations that are compatible with at least one possible evolution (i.e. $A(x) \neq \emptyset$ for all $x \in X_t$), which may change in time. For example, the set of possible configurations for a system under dissipative processes will shrink as they converge to equilibrium. In quantum mechanics, the possible configurations after a measurement are restricted to the eigenstates, and a different choice of measurement will lead to a different process with different possible evolutions. Secondly, the best possible description may be broader than the full state of the system (e.g. when external interference forces us to give only a statistical account within our $\Delta t$) or narrower (e.g. when there are known correlations to the other degrees of freedom of the environment or of the system itself). In other words: the snapshots depend on the process and on other systems within the same process. These correlations prevents the snapshots to be proper states of a system, since they are not descriptions about and only about the system.

\textbf{Determinism}. We say a process is \textbf{deterministic} over the system domains $\{\tdomain_t\}_{t \in T}$ if given the snapshot at one time we can always predict the snapshot at all future times. Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_1 \in X_{t_1} \subset \tdomain_{t_1}$ we can find an $x_2 \in X_{t_2} \subset \tdomain_{t_2}$ such that $x_1 \narrower x_2$. Intuitively, all the evolutions that pass through $x_1$ will also pass through $x_2$. If $x(t)$ is the snapshot trajectory in time, we have $A(x(t)) \subseteq A(x(t + \Delta t))$: the set of compatible evolutions must stay the same or become larger. This also means that the process entropy associated to a snapshot cannot decrease during deterministic evolution:
\begin{equation}\label{entropy_determinism}
	S_{\stmt[u]}(x(t)) \leq S_{\stmt[u]}(x(t + \Delta t))
\end{equation}
since $x(t)\narrower x(t+\Delta t)$ means $x(t) \finer x(t+\Delta t)$ and $\mu_{\stmt[u]}(x(t)) \leq \mu_{\stmt[u]}(x(t+\Delta t))$.

\textbf{Law of evolution}. Deterministic processes are exactly the ones for which a law of evolution can be written. For each $x_1 \in X_1$, the final snapshot $x_2 \in X_2$ must be unique as all snapshots in $X_2$ are incompatible with each other. Therefore we can write a function $f : X_1 \to X_2$ such that $x_1 \narrower f(x_1) \equiv x_2$ that describes that particular step in the process. We call $f$ the \textbf{law of evolution}. The law of evolution fully characterize the deterministic process.

\textbf{Scientific privileging of deterministic processes}. Deterministic processes play a special role in science because, insofar we want to have theories that make prediction, we are going necessarily focus on deterministic processes. This privileging is not due to nature per se: it is due to the practice of scientific investigation. We setup initial conditions and measure results, not the other way around. The time asymmetry is built into the practice of science as a fundamental trait. As such, it is likely to be taken as starting point, and cannot further be explained through scientific investigation itself.

\textbf{Reversibility}. Conversely, we say a process is \textbf{reversible} over the system domains $\{\tdomain_t\}_{t \in T}$ if given the snapshot at one time we can always reconstruct the snapshot at all past times. Formally, let $t_1, t_2 \in T$ such that $t_1 \leq t_2$, for any $x_2 \in X_{t_2} \subset \tdomain_{t_2}$ we can find an $x_1 \in X_{t_1} \subset \tdomain_{t_1}$ such that $x_2 \narrower x_1$. In this case, all the evolutions that pass through $x_2$ must have passed through $x_1$. If $x(t)$ is the snapshot trajectory in time, we have $A(x(t)) \subseteq A(x(t - \Delta t))$: the set of compatible evolutions must stay the same or become smaller.  This also means that the process entropy associated to a snapshot cannot increase during reversible evolution:
\begin{equation}\label{entropy_reversibility}
	S_{\stmt[u]}(x(t)) \leq S_{\stmt[u]}(x(t - \Delta t))
\end{equation}
since $x(t)\narrower x(t-\Delta t)$ means $x(t) \finer x(t-\Delta t)$ and $\mu_{\stmt[u]}(x(t)) \leq \mu_{\stmt[u]}(x(t-\Delta t))$.

\textbf{Law of inverse evolution}. In a dual fashion, reversible processes allow us to write a law of inverse evolution. For each $x_2 \in X_2$, the initial snapshot $x_1 \in X_1$ must be unique as all snapshots in $X_1$ are incompatible with each other. Therefore we can write a function $g : X_2 \to X_1$ such that $x_2 \narrower g(x_2) \equiv x_1$ that describes that particular inverse step in the process. We call $g$ the \textbf{law of inverse evolution}.

\textbf{Determinism and reversibility}. A deterministic and reversible process, then, will mean $x_1$ and $x_2$ are equivalent and that  $A(x(t)) = A(x(t + \Delta t))$: the set of evolutions remains the same. This also means that the process entropy remains conserved:
\begin{equation}\label{entropy_detrev}
	S_{\stmt[u]}(x(t)) = S_{\stmt[u]}(x(t - \Delta t))
\end{equation}

The fact that past and future snapshots are equivalent $x(t) \equiv x(t + \Delta t)$ does not mean they are the same statement in terms of the description at each time: it means there is an if-and-only-if relationship between the descriptions at different times. For example, \statement{at time $t$ the position is $q$ and the velocity is $v$} if and only if \statement{at time $t + \Delta t$ the position is $q + v \Delta t$ and the velocity is $v$}. The differential equation $\dot{v} = 0$ is a short hand for all such relationships.

\textbf{Entropy increase and irreversibility}. According to these definitions, \textbf{entropy cannot decrease in deterministic processes}. Additionally, \textbf{reversible process (that are also deterministic) will conserve entropy}. For non-deterministic processes this may not be the case, therefore this is not an absolute law. However, insofar we want to write theories that give predictions, the result will apply. Therefore it is not that entropy never decreases, it never decreases for the processes we are interested in: the ones where we are able to properly define systems and make predictions. We can say that it is a \textbf{fundamental law of scientific theories but not a fundamental law of nature}.

\textbf{Reversibility vs retrodictability}. If the process is deterministic and reversible we have that the inverse of law of evolution $f^{-1}=g$ is equal to the law of inverse evolution. It is possible for $f$ to be invertible even if the system is \emph{not} reversible. In fact, we would have $x_1 \equiv f^{-1}(x_2) \narrower x_2$, which is not the same as $x_2 \narrower g(x_2) \equiv x_1$: narrowness is in the opposite direction. In other words, \textbf{an invertible deterministic system is not necessarily reversible}. For example, consider a damped harmonic oscillator: the dynamics is in principle invertible (we can reconstruct the trajectory) but it is not reversible (as the evolutions bunch together, finite precision knowledge of the initial conditions gives us more information than the same finite precision knowledge of the final conditions). These types of distinctions are crucial and they emerge naturally within the framework. It can be shows that this is possible only for systems identified by continuous variables, and not when all variables are discrete.

TODO: picture with deterministic evolution merging "streams" of evolutions, and detrev evolutions not merging

\begin{table}[h!]
	\centering
	\begin{tabular}{|c|c|p{6cm}|}
		\hline 
		Symbol & Name & Meaning \\ 
		\hline 
		$t \in T$ & Time parameter & Time is treated as a real valued parameter within the range $T \subseteq \mathbb{R}$ \\ 
		\hline 
		$\tdomain_t \subset \tdomain[P]$ & System domain & The set of descriptions at a given time \\ 
		\hline 
		$X_t \subset \tdomain_t$ & Possible snapshots & The set of descriptions that give a full account at time $t$; they correspond to the narrowest statements within $\tdomain_t$ \\ 
		\hline 
		$x(t) \narrower x(t + \Delta t)$ & Determinism & Given the snapshot at one time we can predict the snapshot at future times \\ 
		\hline 
		$x(t) \narrower x(t - \Delta t)$ & Reversibility & Given the snapshot at one time we can reconstruct the snapshot at past times \\ 
		\hline 
	\end{tabular} 
	\caption{System domain, snapshots, determinism and reversibility}
	\label{table:states}
\end{table}

\subsection{Future: States and state spaces}

In other words, the state space of the system (i.e. the set of all possible complete descriptions of the system and only about that system, independently of the process at hand) is something we need to construct. It is an endpoint, not a starting point. This construction comes with a set of assumptions about the system which are at the foundations of thermodynamics specifically and physics more in general, and they need to be made explicit. Best descriptions of a system at a time, instead, do not require any additional assumptions and therefore are more general concepts.

\textbf{States and independence}. It should be noted that the same description for the same system in different conditions (i.e. in different processes or at different times in the same process) will not in general correspond to the same count of evolutions. Consider a dynamical system with an attractor, such as a damped harmonic oscillator. As time proceeds, more and more evolutions will converge to the attractor. The same finite region around the attractor will correspond to more and more evolutions. In general, a process may introduce correlations between different variables which influences the value of the process entropy for a specific description.

If two systems are independent, however, the evolution of each system does not constrain the evolution of the others. In this case, the count of evolution, and therefore the entropy, of a particular system is only a function of the description of the independent system. The count of evolution factorizes and the entropy becomes additive.

What is crucial to understand here is that \textbf{independence is a requirement for the definition of a system}: we can talk about a system only insofar we have at our disposal a process under which the system is independent for the rest, such that the description of the system is truly a description of the system, with no correlations. We can speak of a chair as its own system because we have circumstances in which we can interact with the chair in a way that does not affect the other chairs or the tables.

Philosophers talk about description of reality that is joint-carving meaning that reality already contains divisions, prior to our investigation. This is an epistemological version of it: nature gives us access to processes for which some systems are independent from the rest. It is the accessibility of these processes that allows us to define systems, without them we cannot divide nature into parts. That is, if an object is really made of parts but the process driving their evolution is such that the correlations can never be eliminated in a controlled fashion, the parts cannot be defined as independent systems.

\section{Current attempt}

The math.

\section{Examples}

\section{Insightful failures}

\bibliography{bibliography}

\end{document}

### 2021-05-11 Overviews/Informational granularity/InformationGranularity.tex
InfGran: Review with Christine up to 3.2 included

@@
 \begin{description}
-    \item[Normalization.] The value for the unit is one,  $\mu_{\stmt[u]}(\stmt[u]) = 1$.
-    \item[Monotonicity.] It is a monotonic function with respect to fineness; if $\stmt_1 \finer \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
-    \item[Finite additivity.] If $\stmt_1 \ncomp \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$.
+    \item[Normalization] the value for the unit is one,  $\mu_{\stmt[u]}(\stmt[u]) = 1$.
+    \item[Monotonicity] it is a monotonic function with respect to fineness; if $\stmt_1 \finer \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
+    \item[Finite additivity] if $\stmt_1 \ncomp \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$.
 \end{description}
@@
 \textbf{Equigranular possibilities}.
-If a possibility is taken as a unit, then the equigranularity of possibilities is a sufficient condition and it leads to the counting measure on the possibilities.
+If a possibility is taken as a unit, then the equigranularity of possibilities is a sufficient condition for a unique measure and it leads to the counting measure on the possibilities.
 This is strictly not necessary: we could have a domain with 5 possibilities $\{x_1, x_2, x_3, y_1, y_2\}$ such that $x_1 \eqgran x_2 \eqgran x_3$, $y_1 \eqgran y_2$ and $x_1 \OR x_2 \OR x_3 \eqgran y_1 \OR y_2$.
@@
 This is tightly linked with a notion of infinitesimal (i.e.
-the ordering over the level of granularity is non-archemedian, while the real numbers are) which we will see in the next subsection.
-The address this collapse in a controlled way, the idea is to proceed in the following way:
+the ordering over the level of granularity is not Archimedean, while the real numbers are) which we will see in the next subsection.
+To address this collapse in a controlled way, the idea is to proceed in the following way:
 \begin{enumerate}
@@
     \item show that $\lequ$ is order isomorphic to the reals
-    \item show that we can pick a particular isomorphism that respects linear addittivity.
+    \item show that we can pick a particular isomorphism that respects linear additivity.
 \end{enumerate}
@@
 
-\textbf{Overview}.
+\textbf{Archimedean property}.
 The measure zero sets are related to the notion of infinitesimals and the Archimedean property.
 The Archimedean property is typically discussed in terms of linearly ordered groups and fields.
-Roughly speaking, if we pick an element $y$ an element $x$ is infinitesimal with respect to $y$ if $\sum\limits_{i=1}^n x < y$ for any $n$.
+Roughly speaking, if we pick an element $y$, an element $x$ is infinitesimal with respect to $y$ if $\sum\limits_{i=1}^n x < y$ for any $n$.
 The group is Archimedean if no element is infinitesimal with respect to another.
@@
 The (positive) real numbers are Archimedean.
-If we take a lattice of sets and use disjoint union as addition, the lattice is not Archimedean.
-A measure will therefore need a non-Archimedean structure (i.e the sets) into an Archimedean one (i.e.
+On the other hand, consider the lattice of subsets of the real numbers.
+If we use disjoint union as addition, the lattice is not Archimedean: the singletons will be infinitesimals with respect to any finite interval.
+A measure will therefore need to map a non-Archimedean structure (i.e the sets) into an Archimedean one (i.e.
 the positive reals).
 Infinitesimals therefore need to be mapped to zero.
-Note that for the measure to be countably additive, we need to have that the countable sum of infinitesimals is still infinitesimals.
+Note that for the measure to be countably additive, we need to have that the countable sum of infinitesimals is still infinitesimal.
 This will not be true in general: a set with finitely many integers is infinitesimal with respect to the set of all integers; yet the set of all integers is the disjoint union of countably many finite sets of integers.
@@
 So, the ordering he has in mind is fineness for us.
-And then he says that this means that the 4 postulates give a non-Archimedean structure.
+And then he says that this means that his 4 postulates give a non-Archimedean structure.
 But you can make it Archimedean by ignoring the probabilities that are ``infinitesimally small'': those that multiplied by a number, however big, will never become the certainty.
-On the other hand, Villegas\cite{villegas} starts already with an ordering where the infinitesimals are equal to the empty set.
+On the other hand, Villegas\cite{villegas} starts already with a preorder where the infinitesimals are equal to the empty set.
 This maps to $\lequ$ defined later, where the comparison is done up to infinitesimals.
@@
 \textbf{Definition of infinitesimals}.
-The first issue is to find the correct definition to extend this concept on ordering.
+The first issue is to find the correct definition to extend this concept to partially ordered sets.
 The addition discussed above (i.e.
-the disjoint union of statements) should play the role of the group operation.
-Conceptually, a statement $\stmt$ is infinitesimal with respect to another $\stmt[u]$ if for any $n$ we can find a finite set of equigranular incompatible statements whose disjunction is finer than $\stmt[u]$.
-That is, we need infinitely many statements equigranular to $\stmt$ to cover a statement of the same granularity than $\stmt[u]$.
+the disjoint disjunction of statements) should play the role of the group operation.
+Conceptually, a statement $\stmt$ is infinitesimal with respect to another $\stmt[u]$ if for any $n$ we can find a finite set of $n$ equigranular incompatible statements whose disjunction is finer than $\stmt[u]$.
+That is, we need infinitely many statements equigranular to $\stmt$ to cover a statement of the same granularity as $\stmt[u]$.
 
@@
 
-\textbf{Definition of infinitesimals}.
+\textbf{Countable and uncountable infinitesimals}.
 Assuming those issues are solved, infinitesimals come in two types: countable and uncountable.
-A countable infinitesimal will be such that the disjoint conjunction of countably many equigranular statements may be bigger than the unit.
+A countable infinitesimal will be such that the disjoint disjunction of countably many equigranular statements may be bigger than the unit.
 This is the case where, in the end, the measure will only be finitely additive: infinitely many zero measure elements will sum to a finite element.
-If all infinitesimal with respect to the unit are uncountable, instead, countably many zero measure elements will still form a zero measure element.
+If all infinitesimals with respect to the unit are uncountable, instead, countably many zero measure elements will still form a zero measure element.
 
@@
 All sets of measure zero are uncountable infinitesimals with respect to the unit: since the measure is countably additive, any disjoint union of countably many measure zero sets is going to be measure zero, therefore it will not be as big as the unit.
-Naturally, if we take a uncountably many sets (i.e.
+Naturally, if we take uncountably many sets (i.e.
 all the singletons) we can cover the whole unit.
 Conversely, the full real line can be covered by countably many sets of unit size.
-While the unit interval is infinitesimally small with the respect with the whole line, it is a countable infinitesimal.
+While the unit interval is infinitesimally small with the respect to the whole line, it is a countable infinitesimal.
 This definition of ``finiteness'' may map to physical requirements much better than, for example, the topological notion of compactness.
@@
 The next step is to understand how the properties of $\lequ$ would compare to the requirements set by, for example Villegas \cite{villegas}.
-In the context of comparative probability, there are theorems that allow to go from a preorder to a probability measure.
+In the context of comparative probability, there are theorems that allow one to go from a preorder to a probability measure.
 The idea would be to reuse, as much as possible, the ideas from that literature.
 Once we pick a unit $\stmt[u]$, we can restrict ourselves to all statements narrower than $\stmt[u]$.
-At that point, we are looking for has all the mathematical properties of a probability measure as the biggest element, $\stmt[u]$, has to have measure one.
-
-\textbf{Countable addittivity}.
+At that point, what we have has all the mathematical properties of a probability measure, since the biggest element, $\stmt[u]$, has to have measure one.
+
+\textbf{Countable additivity}.
 One issue will be the problem of countable infinitesimals, which are incompatible with a countable measure.
-Therefore one of the assumptions in comparative probability will effectively exclude the case.
+Therefore one of the assumptions in comparative probability will effectively exclude this case.
 We need to understand how that works, so that we can adapt the proofs and allow for only finitely additive measures.
-The hope is that if we pick a unit for which all infinitesimals are uncountable, then one can demonstrate the requirements for the previous theorems
+The hope is that if we pick a unit for which all infinitesimals are uncountable, then one can demonstrate the requirements for the previous theorems.
 
@@
 It is not clear how to recover this set.
-It may be that the equivalence classes of $\equ$ will make our job easier: if two infinitesimal are not comparable in fineness, they will still be merged together.
+It may be that the equivalence classes of $\equ$ will make our job easier: all infinitesimals will end up in the same equivalence class as the impossibility, regardless of their comparability in fineness.
 Therefore incomparable statements may be comparable in $\lequ$ if they can be compared up to an infinitesimal.
 
-The goal here is to solve the above problems to be find the prerequisites for the following proposition: given a statement $\stmt[u] \in \tdomain$ we can define a maximal set of mutually comparable statements in $\lequ$ that contains $\stmt[u]$, upon which we can define a unique $\mu_{\stmt[u]} : \tdomain_{\stmt[u]} \to \mathbb{R}$ such that:
+The goal here is to solve the above problems to find the prerequisites for the following proposition.
+Given a statement $\stmt[u] \in \tdomain$ we can define a maximal set of mutually comparable statements in $\lequ$ that contains $\stmt[u]$, upon which we can define a unique $\mu_{\stmt[u]} : \tdomain_{\stmt[u]} \to \mathbb{R}$ such that:
 \begin{description}
-    \item[Normalization.] The value for the unit is one,  $\mu_{\stmt[u]}(\stmt[u]) = 1$.
-    \item[Order isomorphism.] $\stmt_1 \lequ \stmt_2$ if and only if $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
-    \item[Finite additivity.] If $\stmt_1 \ncomp \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$.
+	\item[Normalization] the value for the unit is one,  $\mu_{\stmt[u]}(\stmt[u]) = 1$.
+    \item[Order isomorphism] $\stmt_1 \lequ \stmt_2$ if and only if $\mu_{\stmt[u]}(\stmt_1) \leq \mu_{\stmt[u]}(\stmt_2)$.
+    \item[Finite additivity] if $\stmt_1 \ncomp \stmt_2$ then $\mu_{\stmt[u]}(\stmt_1 \OR \stmt_2) = \mu_{\stmt[u]}(\stmt_1) + \mu_{\stmt[u]}(\stmt_2)$.
 \end{description}
-Moreover, if $\stmt[u]$ has no countable infinitesimals, $\mu_{\stmt[u]}$ is countably additive.
+Moreover, if $\stmt[u]$ has no countable infinitesimals (or other more restrictive conditions), $\mu_{\stmt[u]}$ is countably additive.
 
@@
 
-One issue will be, again, proving the consistent of comparability with different units since all the domains $\tdomain_{\stmt[u]}$ do not contain all the statements.
+One issue will be, again, proving the consistency of comparability with different units since all the domains $\tdomain_{\stmt[u]}$ do not contain all the statements.
 
@@
     \begin{enumerate}
-        \item if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \sfiner \stmt_2$ if and only if $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$ (these two tell us that offset monotonicity can be extend to strict fineness and equigranularity)
+        \item if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \sfiner \stmt_2$ if and only if $\stmt \OR \stmt_1 \sfiner \stmt \OR \stmt_2$ (these two tell us that offset monotonicity can be extended to strict fineness and equigranularity)
         \item if $\stmt \ncomp \stmt_1$ and $\stmt \ncomp \stmt_2$ then $\stmt_1 \eqgran \stmt_2$ if and only if $\stmt \OR \stmt_1 \eqgran \stmt \OR \stmt_2$
@@
         \item if $\stmt_1 \sfiner \stmt_2$ then $\NOT \stmt_2 \sfiner \NOT \stmt_1$
-        \item if $\certainty \finer \stmt$ then $\certainty \narrower \stmt$ and therefore $\stmt \equiv \certainty$ (these two tells us that fineness can recover narrowness for certainties and impossibilities)
+        \item if $\certainty \finer \stmt$ then $\certainty \narrower \stmt$ and therefore $\stmt \equiv \certainty$ (these two tell us that fineness can recover narrowness for certainties and impossibilities)
         \item if $\stmt \finer \impossibility$ then $\stmt \narrower \impossibility$ and therefore $\stmt \equiv \impossibility$
@@
     \item (monotony as in Villegas) if $\stmt[v]_1 \ncomp \stmt[v]_2$, then $\stmt_1 \finer \stmt[v]_1$ and $\stmt_2 \finer \stmt[v]_2$ imply  $\stmt_1 \OR \stmt_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
-    \item if $\stmt_1 \ncomp \stmt_2$, $\stmt[v]_1 \ncomp \stmt[v]_2$ and $\stmt_2 \eqgran \stmt[v]_2$, then $\stmt_1 \finer \stmt[v]_1$ if and only if $\stmt_1 \OR \stmt_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
 \end{enumerate}
@@
 
-Since all $\{ \stmt[w]_i \}$ are pair-wise incompatible, offset monotonicity allows to remove or add the same statement on both sides of a fineness relationship. Therefore
+Since all $\{ \stmt[w]_i \}$ are pair-wise incompatible, offset monotonicity allows one to remove or add the same statement on both sides of a fineness relationship. Therefore
 \begin{itemize}
@@
     \item $\stmt[v]_2 \coarser \stmt[s]_2 \implies  \stmt[w]_4 \OR \stmt[w]_5 \OR \stmt[w]_6 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_4 \implies \stmt[w]_5 \OR \stmt[w]_6 \coarser \stmt[w]_2 \OR \stmt[w]_3 \implies \stmt[w]_5 \OR \stmt[w]_6 \OR \stmt[w]_7 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_7$
-    \item $\stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_5 \coarser \stmt[w]_5 \OR \stmt[w]_6 \OR \stmt[w]_7 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_7 \implies \stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_5 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_7 \implies \stmt[w]_1 \OR \stmt[w]_5 \coarser \OR \stmt[w]_3 \OR \stmt[w]_7$
+    \item $\stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_5 \coarser \stmt[w]_5 \OR \stmt[w]_6 \OR \stmt[w]_7 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_7 \implies \stmt[w]_1 \OR \stmt[w]_2 \OR \stmt[w]_5 \coarser \stmt[w]_2 \OR \stmt[w]_3 \OR \stmt[w]_7 \implies \stmt[w]_1 \OR \stmt[w]_5 \coarser  \stmt[w]_3 \OR \stmt[w]_7$
 \end{itemize}
@@
 By 1, we have $\stmt[s]_1 \OR \hat{\stmt[s]}_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
-But since $\stmt[s]_1 \OR \hat{\stmt[s]}_2 \equiv \stmt[s]_1 \OR (\stmt[s]_2 \AND \NOT \stmt[s]_1) \equiv (\stmt[s]_1 \OR \stmt[s]_2) \AND (\stmt[s]_1 \OR \NOT \stmt[s]_1) \equiv \stmt[s]_1 \OR \stmt[s]_2$, we have $\stmt[s]_1 \OR \stmt[s]_2 \finer \stmt[v]_1 \OR \stmt[v]_2$..
+But since $\stmt[s]_1 \OR \hat{\stmt[s]}_2 \equiv \stmt[s]_1 \OR (\stmt[s]_2 \AND \NOT \stmt[s]_1) \equiv (\stmt[s]_1 \OR \stmt[s]_2) \AND (\stmt[s]_1 \OR \NOT \stmt[s]_1) \equiv \stmt[s]_1 \OR \stmt[s]_2$, we have $\stmt[s]_1 \OR \stmt[s]_2 \finer \stmt[v]_1 \OR \stmt[v]_2$.
 
@@
 
-	\item if $\stmt[v]_1\ncomp\stmt[v]_2$, $\stmt[v]_2\eqgran\stmt_2$, and $\stmt_1\finer\stmt[v]_1$. Then $\stmt_1\OR\stmt_2\finer\stmt[v]_1\OR\stmt[v]_2$.
+	\item if $\stmt[v]_1\ncomp\stmt[v]_2$, $\stmt[v]_2\eqgran\stmt_2$, and $\stmt_1\finer\stmt[v]_1$, then $\stmt_1\OR\stmt_2\finer\stmt[v]_1\OR\stmt[v]_2$.
 \end{itemize}
@@
 We call \textbf{granularity level} an equivalence class of equigranularity.
-Let $(\cdot)_{/\eqgran} : \tdomain \to \tdomain_{/\eqgran}$ the map from a statement to its granularity level, which we also note $\dot{\stmt} = {\stmt}_{/\eqgran}$.
+Let $(\cdot)_{/\eqgran} : \tdomain \to \tdomain_{/\eqgran}$ be the map from a statement to its granularity level, which we also note $\dot{\stmt} = {\stmt}_{/\eqgran}$.
 We say $\dot{\stmt}$ is the granularity level of $\stmt$.
@@
 Suppose $\stmt_1 \narrower \stmt_2$ then by \ref{fineness_properties}.3 $\stmt_1 \finer \stmt_2$ and therefore ${\stmt_1}_{/\eqgran} \finer {\stmt_2}_{/\eqgran}$.
-Then $(\cdot)_{/\eqgran}$ is an order isomorphism.
+Then $(\cdot)_{/\eqgran}$ is an order homomorphism.
 \end{proof}
@@
     Regardless of whether $\tdomain_{/\eqgran}$ is a lattice, $(\cdot)_{/\eqgran}$ can never be a lattice homomorphism.
-It will be crucial that two equigranular statements can be incompatible, and this in contradiction with $(\cdot)_{/\eqgran}$ being a lattice homomorphism.
+It will be crucial that two equigranular statements can be incompatible, and this is in contradiction with $(\cdot)_{/\eqgran}$ being a lattice homomorphism.
 Let us note $\dot{\stmt} = {\stmt}_{/\eqgran}$ the level of granularity of $\stmt$.
@@
     Let $\tdomain$ be uniform.
-Let $\stmt, \stmt_1 \in \tdomain$ such that $\stmt \sfiner \stmt_1$.
-Then there exist $\stmt_2 \in \tdomain$ such that $\stmt_2 \narrower \stmt_1$ and $\stmt \eqgran \stmt_2$.
+Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \sfiner \stmt_2$.
+Then there exists $\stmt \in \tdomain$ such that $\stmt \narrower \stmt_2$ and $\stmt \eqgran \stmt_1$.
 \end{desid}
@@
 \begin{defn}\label{infinitesimals}
-	Let $\stmt[u],\stmt[s] \in \tdomain$.
+	Let $\stmt[u],\stmt[s] \in \tdomain$ such that $\stmt[u] \neq \impossibility$.
 Then $\stmt$ is \textbf{infinitesimal} with respect to $\stmt[u]$ if for any natural number $n$ we can find a set of $n$ statements $\{ \stmt_i \}_{i=1}^n$ such that:
@@
 	\end{enumerate}
-	In words, $\stmt$ is infinitesimal with respect to $\stmt[u]$ if for any number, there is a disjunction of that many broader mutually incompatible statements $\stmt_i$ is finer than $\stmt[u]$.
+	In words, $\stmt$ is infinitesimal with respect to $\stmt[u]$ if for any number, there is a disjunction of that many broader mutually incompatible statements $\stmt_i$ that is finer than $\stmt[u]$.
 \end{defn}
@@
 
-TODO[Mark]: check the proof
 \begin{prop}
@@
 By \ref{fineness_monotony} we have $\bigOR_{i=1}^n \stmt_i \finer \bigOR_{i=1}^n \stmt[t]_i$ and therefore $\stmt[u] \finer \bigOR_{i=1}^n \stmt_i \finer \bigOR_{i=1}^n \stmt[t]_i$.
-Therefore we cannot find a set of $n$ statement that satisfy the definition of infinitesimal and therefore $\stmt$ is not an infinitesimal with respect to $\stmt[u]$.
+Therefore we cannot find a set of $n$ statements that satisfy the definition of infinitesimal and therefore $\stmt$ is not an infinitesimal with respect to $\stmt[u]$.
 \end{proof}
@@
 Without loss of generality, suppose $\stmt[v] \finer \stmt[w]$.
-Let $n>0$.
+Let $n>0$ be a natural number.
 Let $\{\stmt[w]_i\}_{i=1}^{2n}$ be a collection realizing infinitesimality for $2n$ for $\stmt[w]$.

### 2021-05-14 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: more work on cardinals and ordinals

@@
 
+\def\ordinals{\textbf{ORD}}
+\def\cardinals{\textbf{CRD}}
+
+\def\ordless{\prec}
+\def\ordleq{\preceq}
+\def\ordeq{\sim}
+\def\ordgeq{\succeq}
+\def\ordgtr{\succ}
+
+\def\crdless{\sqsubset}
+\def\crdleq{\hookrightarrow}
+\def\crdeq{\leftrightarrow}
+\def\crdgeq{\hookleftarrow}
+\def\crdgtr{\sqsupset}
+
 \title{Bare minimum: set theory}
@@
 \end{abstract}
+
+Ordinal: $\ordless \ordleq \ordeq \ordgeq \ordgtr$
+
+Cardinal: $\crdless \crdleq \crdeq \crdgeq \crdgtr$
 
@@
 	 The \textbf{empty set}
-	 \marginleft{Common sets: $\emptyset$, $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$, $\mathbb{C}$}
-	  $\emptyset$ is the set with no elements.
+	 \marginleft{Common sets: $\emptyset$, $\{a\}$, $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$, $\mathbb{C}$}
+	  $\emptyset$ is the set with no elements. A \textbf{singleton} is the set of a single value $\{a\}$.
 	 The \textbf{set of natural numbers} $\mathbb{N} = \{ 0, 1, 2, ...
@@
 \begin{defn}[Power set]
-	Given a set $A$, its \textbf{power set}, \marginleft{Power set: $2^{A}$, $\mathcal{P}(A)$} noted $\mathcal{P}(A)$ or $2^{A}$, is the set of all subsets of $A$.
+	Given a set $A$, its \textbf{power set}, \marginleft{Power set:\\ $\mathcal{P}(A)$, $2^{A}$} noted $\mathcal{P}(A)$ or $2^{A}$, is the set of all subsets of $A$.
 That is, $\mathcal{P}(A) = \{B \, | \, B \subset A \}$.
@@
 
-	Let $X \subseteq \mathcal{P}(A)$ be a set of subsets of $A$. The operations of union $\bigcup\limits_{Y \in X} Y$ and intersection $\bigcap\limits_{Y \in X} Y$ can be extend to span all elements.
+	Let $\mathcal{A} \subseteq \mathcal{P}(A)$ be a set of subsets of $A$.
+The operations of union $\bigcup\limits_{A \in \mathcal{A}} A$ and intersection $\bigcap\limits_{A \in \mathcal{A}} A$ can be extend to span all elements.
 \end{defn}
 
-We give the following informal definition. Given two elements $a$ and $b$, the \textbf{ordered pair} $(a, b)$ specifies two objects in that order. That is $(a, b) \neq (b, a)$ and $(a, a) \neq \{a\}$.
+\begin{defn}[Ordered pair]
+	Given two \marginleft{Ordered pair: $(a, b)$} elements $a$ and $b$, the \textbf{ordered pair} $(a, b)$ specifies two objects in that order.
+That is $(a, b) \neq (b, a)$.
+This can be constructed setting $(a, b) = \{ \{a\}, \{a, b\}\}$.
+\end{defn}
 
@@
 	Given two sets \marginleft{Cartesian product: $\times$}  $A$ and $B$, the \textbf{Cartesian product} $A \times B = \{ (a,b) \, | \, \forall a \in A, b \in B \}$ is the set of all possible ordered pairs between the elements of $A$ and $B$.
+\end{defn}
+
+\begin{defn}[Disjoint union]
+	Let $A$ and $B$ be two sets. The disjoint union $A \sqcup B$ is the union where the elements of the sets are always treated as distinct. That is, $A \sqcup B = \{A\} \times A \cup \{B\} \times B$.
 \end{defn}
@@
 
+\begin{defn}
+	Let $f : A \to B$ be a function.
+This induces a relation $f : \mathcal{P}(A) \to \mathcal{P}(B)$ that associates each subset of $A$ to its image and a relation $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ that associates each subset of $B$ with its reverse image.
+\end{defn}
+
+\begin{defn}[Level set]
+	A level set \marginleft{Level set} of a function $f : A \to B$ is a set of where the function takes a given value $b$. In other words the reverse image $f^{-1}(\{c\})$ a singleton.
+\end{defn}
+
+\begin{prop}
+	The induces relations $f : \mathcal{P}(A) \to \mathcal{P}(B)$ and $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ are functions and satisfy the following properties:
+	\begin{enumerate}
+		\item $f(\bigcup\limits_{A \in \mathcal{A}} A) = \bigcup\limits_{A \in \mathcal{A}} f(A)$
+		\item $f^{-1}(\bigcup\limits_{A \in \mathcal{A}} A) = \bigcup\limits_{A \in \mathcal{A}} f^{-1}(A)$
+		\item $f^{-1}(\bigcap\limits_{A \in \mathcal{A}} A) = \bigcap\limits_{A \in \mathcal{A}} f^{-1}(A)$
+	\end{enumerate}
+\end{prop}
+
+\begin{remark}
+	With abuse of notation, $f$ and$f^{-1}$ indicate both the function between elements and sets.
+The image of the intersection is not in general the intersection of the image.
+Note how in both topology and measure theory continuity and measurability are defined based on the reverse images.
+\end{remark}
+
 \begin{defn}[Identity function]
@@
 \end{prop}
+
+\begin{defn}[Restriction]
+	Let $f : A  \to B$ \marginleft{Restriction: $f|_C$} be a function and let $C \subseteq B$. Then the \textbf{restriction of $f$ over $C$} is the function $f|_C : C \to B$ such that $f|_C(c) = f(c)$.
+\end{defn}
 
@@
 \end{defn}
+
+\begin{prop}
+	Let $2 = \{0, 1\}$ denote a set with two elements. Then $\mathcal{P}(A)$ and $2^A$ are in one-to-one correspondence.
+\end{prop}
 
@@
 	\begin{enumerate}
-		\item $\bigcup\limits_{X \in P} X = A$
-		\item $X \cap Y = \emptyset$ for all $X, Y \in P$ where $X \neq Y$
+		\item $\bigcup\limits_{B \in P} B = A$
+		\item $B \cap C = \emptyset$ for all $B, C \in P$ where $B \neq C$
 	\end{enumerate}
@@
 
+\begin{defn}
+	Given a function $f : A \to B$ the \textbf{equivalence relation determined by $f$} is the relation $R_f = \{ (a_1, a_2) \, | \, f(a_1) = f(a_2) \}$.
+\end{defn}
+
+\begin{prop}[Function canonical decomposition]
+	Every function $f : A \to B$ can be written as the composition $f = t \circ s \circ r$ of the following three functions:
+	\begin{description}
+		\item[$r : A \to A_{/R_f}$] defined as $a \mapsto [a]_{R_f}$; this function is surjective
+		\item[$s : A_{/R_f} \to f(A)$] defined as $[a]_{R_f} \mapsto f(a)$; this function is bijective
+		\item[$t : f(A) \to B$] defined as $t(b) = b$; this function is injective.
+	\end{description}
+\end{prop}
+
 %\begin{prop}[Set inclusion as ordering]
@@
 \end{remark}
-\begin{defn}
+\begin{defn}\label{defn_successor_set}
 	Let $A$ be a set.
@@
 	\end{itemize}
+	It can be shown that they obey Peano's axioms, another common way to axiomatize the natural numbers and their arithmetic.
+\end{remark}
+
+\subsection{Ordinal numbers}
+
+For infinite sets, cardinals and ordinals are different since sets of the same infinite size can be given different orderings.
+
+\begin{remark}
+	Cardinal numbers (i.e.
+one, two, three, ...) identify the number of elements in a set.
+Ordinal numbers (i.e.
+first, second, third) identify a position within a sequence.
+Well-ordered sets extend the notion of sequence to arbitrary levels of infinity.
+\end{remark}
+
+\begin{defn}[POS]
+	A \textbf{partially ordered set}, \marginleft{POS: $(A, \leq_A)$} or \textbf{POS}, is a tuple $(A, \leq_A)$ where $A$ is a set and $\leq_A$ is a partial order over $A$.
+A \textbf{linearly ordered set} is a POS whose order is linear.
+\end{defn}
+
+\begin{defn}[Initial segment]
+	The \textbf{initial segment} \marginleft{Initial segment} of a POS $(A, \leq_A)$ determined by $a \in A$ is the set of all elements less than $a$. That is, $S_a=\{ x \in A \, | \, x < a \}$.
+\end{defn}
+
+\begin{defn}[Order morphisms]
+	Let $(A, \leq_A)$ \marginleft{Order homo/ \\ isomorphism} and $(B, \leq_B)$ be two POSs.
+An \textbf{order homomorphism} is a function $f : A \to B$ that preserves the ordering.
+That is, $a_1 \leq_A a_2$ implies $f(a_1) \leq_B f(a_2)$.
+An \textbf{order isomorphism} is a bijective order homomorphism whose inverse is also an order homomorphism.
+That is, $b_1 \leq_B b_2$ implies $f^{-1}(b_1) \leq_A f^{-1}(b_2)$.
+\end{defn}
+
+\begin{defn}[Well-order]
+	A \textbf{well-ordered} \marginleft{Well-order} set is a linearly ordered set $(A, \leq_A)$ where each subset has a least element.
+That is, for every $B \subseteq A$, there exists an $a \in B$ such that $a \leq_A b$ for all $b \in B$.
+Let $A$ and $B$ be two well ordered set, we say $A$ \textbf{has lower ordinality than} $B$, noted $A \ordleq B$, if $A$ is order isomorphic to an initial segment of $B$.
+They have \textbf{the same ordinality}, noted $A \ordeq B$, if they are order isomorphic.
+\end{defn}
+
+\begin{prop}[Ordinality]
+	If $A$ and $B$ are two well ordered sets, then either $A \ordleq B$ or $B \ordleq A$. If $\mathcal{A}$ is a collection of well-ordered sets, then one element has least ordinality.
+\end{prop}
+
+\begin{remark}
+	Examples of well-ordered sets:
+	\begin{itemize}
+		\item a finite sequence: $(a_1, a_2, ..., a_n)$
+		\item an infinite sequence: $(a_1, a_2, ..., a_n, ...)$
+		\item an infinite sequence with elements afterwards: $(a_1, a_2, ..., a_n, ..., b_1, b_2, ... , b_6)$
+	\end{itemize}
+	In a well-ordered set, every element always has an immediate successor, but not necessarily an infinite predecessor (e.g. $b_1$ in the third example). 
+\end{remark}
+
+\begin{remark}
+	Ordinal numbers generalize the construction of von Neumann ordinals and cover all possible positions in well-ordered sets.
+\end{remark}
+
+\begin{defn}[Ordinals]
+	An \textbf{ordinal} \marginleft{Ordinals: \ordinals} is a set $A$ with the following properties for all $a, b, c \in A$:
+	\begin{enumerate}
+		\item $a \notin a$
+		\item $a \in b$ implies $b \notin a$
+		\item $a \in b$ and $b \in c$ implies $a \in c$
+		\item $x \in a$ implies $x \in A$
+	\end{enumerate}
+	We note $\ordinals$ as the class of all ordinals.
+\end{defn}
+
+\begin{prop}[Ordinals are well-ordered]
+	Let $a \in \ordinals$ be an ordinal. Then $(a, \in)$ is a well-ordered set.
+\end{prop}
+
+\begin{prop}[Ordinality of a well-ordered set]
+	For every \marginleft{$\langle A \rangle$} well-ordered set $A$ there is a unique $\langle A \rangle \in \cardinals$ such that $a \ordeq A$.
+We say $\langle A \rangle$ is the \textbf{ordinality of $A$}.
+The natural numbers are the initial segment of the ordinals.
+$\omega$ is the ordinal associated with the natural numbers.
+\end{prop}
+
+\begin{remark}
+	As an example, consider the von Neumann ordinal $3 = (0, 1, 2)$ as a well-ordered set.
+We can see it as a sequence of three element, which means the last is the third.
+Similarly, the ordinality of the set $(a_1, a_2, a_3)$ is $3 = (0, 1, 2)$ as they define well-ordered set with the same ordering.
+The definitions are such that these ideas extend to infinite sets.
+\end{remark}
+
+\begin{defn}[Ordinal addition]
+	Let $(A, \leq_A)$ \marginleft{Ordinal addition: $\langle A \rangle + \langle B \rangle$} and $(B, \leq_B)$ be two well ordered sets. Let $(A \sqcup B, \leq)$ be the POS where:
+	\begin{itemize}
+		\item $(A, a_1) \leq (A, a_2)$ if $a_1 \leq_A a_2$
+		\item $(B, b_2) \leq (B, b_2)$ if $b_1 \leq_B b_2$
+		\item $(A, a) \leq (B, b)$ for all $a \in A$ and $b \in B$.  	
+	\end{itemize}
+	Then $(A \sqcup B, \leq)$ is well-ordered and the ordinal $\langle A \sqcup B \rangle$ is fully determined by $\langle A \rangle$ and $\langle B \rangle$.
+Therefore we can define the \textbf{ordinal addition} $\langle A \rangle + \langle B \rangle =  \langle A \sqcup B \rangle$.
+\end{defn}
+
+%\begin{prop}[Ordinal multiplication]The addition has the following properties:
+%	\begin{enumerate}
+%		\item associative: $\langle A \rangle + ( \langle B \rangle + \langle C \rangle) = (\langle A \rangle + \langle B \rangle) + \langle C \rangle$
+%		\item $\langle 0 \rangle = \langle \emptyset \rangle$ is the identity: $\langle A \rangle + \langle 0 \rangle = \langle 0 \rangle + \langle A \rangle = \langle A \rangle$
+%		\item not commutative in general: $\langle A \rangle + \langle B \rangle$ not necessarily the same as $\langle B \rangle + \langle A \rangle$
+%		\item reduces to arithmetic addition on the natural numbers
+%	\end{enumerate}
+%\end{prop}
+
+\begin{remark}
+	Addition in non-commutative (i.e.
+$\langle A \rangle + \langle B \rangle \neq \langle B \rangle + \langle A \rangle$) for infinite ordinals.
+For example $(a_1, a_2, ..., a_n, ...) \sqcup (b_1, b_2, b_3) = (a_1, a_2, ..., a_n, ..., b_1, b_2, b_3)$ is not the same as $(b_1, b_2, b_3) \sqcup (a_1, a_2, ..., a_n, ...) = (b_1, b_2, b_3, a_1, a_2, ..., a_n, ...) \simeq (c_1, c_2, ..., c_n, ...)$.
+Therefore $\omega + 3 \neq 3 + \omega = \omega$.
+\end{remark}
+
+\begin{defn}[Ordinal multiplication]
+	Let $(A, \leq_A)$ \marginleft{Ordinal multiplication: $\langle A \rangle \langle B \rangle$} and $(B, \leq_B)$ be two well ordered sets. Let $(A \times B, \leq)$ be the POS where:
+	\begin{itemize}
+		\item $(a_1, b_1) \leq (a_2, b_2)$ if $b_1 \leq_B b_2$
+		\item $(a_1, b) \leq (a_2, b)$ if $a_1 \leq_A a_2$
+	\end{itemize}
+	Then $(A \times B, \leq)$ is well-ordered and the ordinal $\langle A \times B \rangle$ is fully determined by $\langle A \rangle$ and $\langle B \rangle$.
+Therefore we can define the \textbf{ordinal multiplication} $\langle A \rangle \langle B \rangle =  \langle A \times B \rangle$.
+\end{defn}
+
+%\begin{prop}[Ordinal multiplication]
+%	The multiplication has the following properties:
+%	\begin{enumerate}
+%		\item associative: $\langle A \rangle ( \langle B \rangle \langle C \rangle) = (\langle A \rangle \langle B \rangle) \langle C \rangle$
+%		\item $\langle 1 \rangle = \langle \{ \emptyset \} \rangle$ is the identity: $\langle A \rangle \langle 1 \rangle  = \langle 1 \rangle \langle A \rangle = \langle A \rangle$
+%		\item right distributive over addition: $\langle A \rangle ( \langle B \rangle + \langle C \rangle) = \langle A \rangle \langle B \rangle + \langle A \rangle \langle C \rangle$
+%		\item $\langle A \rangle \langle 0 \rangle  = \langle 0 \rangle \langle A \rangle = \langle 0 \rangle$
+%		\item not commutative in general: $\langle A \rangle + \langle B \rangle$ not necessarily the same as $\langle B \rangle + \langle A \rangle$
+%		\item reduces to arithmetic multiplication on the natural numbers
+%	\end{enumerate}
+% The multiplication is associative and $\langle \emptyset \rangle$ is the identity (i.e.
+$\langle A \rangle + \langle \emptyset \rangle = \langle \emptyset \rangle + \langle A \rangle = \langle A \rangle$).
+The additions is not, in general, commutative.
+%\end{prop}
+
+\begin{remark}
+	Muliplication is non-commutative (i.e.
+$\langle A \rangle \langle B \rangle \neq \langle B \rangle \langle A \rangle$) for infinite ordinals.
+For example $(a_1, a_2, ..., a_n, ...) \times (b_1, b_2) = ((a_1, b_1), (a_2, b_1), ..., (a_n, b_1), ..., \\ (a_1, b_2), (a_2, b_2), ..., (a_n, b_2), ...)$ is not the same as $(b_1, b_2) \times (a_1, a_2, ..., a_n, ...) = ((b_1, a_1), (b_2, a_1), (b_1, a_2), (b_2, a_2), ..., (b_1, a_n), (b_2, a_n), ...) \simeq (c_1, c_2, ..., c_n, ...)$.
+\\ Therefore $\omega 2 \neq 2 \omega = \omega$.
+
+	With addition and multiplication, we can get a sense of how the ordinals work.
+$\omega$ represents a countable sequence.
+$\omega +3$ is a countable sequence followed by three elements.
+$\omega + \omega  = \omega2$ represents a countable sequence followed by a countable sequence.
+$\omega6$ represents a sequence of 6 countable sequences.
+$\omega\omega = \omega^2$ represents a countable sequence of countable sequences.
+$\omega^3$ represents a countable sequence of countable sequences of countable sequences.
+$\omega^\omega$ extends the recursion infinitely many times and so on.
+There is no upper limit to the ordinals.
+\end{remark}
+
+\begin{remark}
+	The cardinality of a set (i.e.
+how many elements it contains) is defined in terms of injection/bijection: a set is smaller than another if it can be injected into it, and it is as big as another if it can be put into a one-to-one correspondence.
+\end{remark}
+
+\begin{defn}[Cardinal comparison]
+	Let $A$ \marginleft{Cardinality} and $B$ be two sets.
+If there exists a injective function $f : A \to B$ then we say $A$ \textbf{has lower cardinality than} $B$, noted $A \crdleq B$.
+If there exists a bijective function we say they have \textbf{the same cardinality}, noted $A \crdeq B$.
+\end{defn}
+
+\begin{prop}[Cardinal comparability of sets]
+	Given two sets $A$ and $B$, either $A \crdleq B$ or $B \crdleq A$ one must have lower cardinality than the other (i.e.
+there exists an injective function from one two the other).
+Additionally, $A \crdeq B$ if and only if $A \crdleq B$ and $B \crdleq A$ (i.e.
+there exists a bijective function between $A$ and $B$ only if there exists an injective function in both direction).
+\end{prop}
+
+\begin{prop}[Well-ordering theorem]
+	Every set can be well-ordered.
+\end{prop}
+
+\begin{remark}
+	The well-ordering theorem allows us to reuse ordinal numbers to quantify cardinality. As two different well-ordered sets may have the same cardinality, we'll take the lowest ordinal between them.
+\end{remark}
+
+\begin{defn}[Cardinal number]
+	A \textbf{cardinal}  \marginleft{Cardinals: $\cardinals$} is an ordinal $a \in \ordinals$ that has lower ordinality of any ordinal that has the same cardinality.
+That is, $a \ordleq b$ for all $b \in \ordinals$ such that $a \crdeq b$.
+The set of all cardinals is noted $\cardinals \subset \ordinals$.
+\end{defn}
+
+\begin{prop}[Cardinality of a set]
+	For every \marginleft{$|A|$} set $A$ there is a unique $|A| \in \cardinals$ such that $a \crdeq A$. We say $|A|$ is the \textbf{cardinality of $A$}.
+\end{prop}
+
+\begin{remark}
 \end{remark}
@@
 
-Syntactic paradoxes two main strategies.
-Zermelo–Fraenkel (ZFC) only talk about sets and restrict set construction to subsets only.
-von Neumann–Bernays–Gödel (NBG) talks about classes and defines sets as those classes that can be elements of classes.
-Set builder notation works on sets, not classes.
-So, if putting a class in a class creates a contradiction, it's because the class is not a set, but a proper class.
+\subsection{Classes, sets and elements}
+
+As some collections cannot \marginright{TODO: diagram with Classes, proper classes and sets.} be put into other collections without creating paradoxes, the resolution of syntactic paradoxes is through the distinction between classes and sets.
+Classes are all those objects that can contain elements.
+Sets are particular classes that can also be elements of other classes.
+Therefore we can have the class of all sets, but we cannot have the set of all sets or the class of all classes.
+
+\subsection{Axiomatization}
+
+There are different approaches to the axiomatization of set theory with different degree of variation.
+One strategy, refered to as Zermelo–Fraenkel (ZFC), only defines sets.
+It avoids paradoxes by restricting set construction to subsets only.
+
+Another strategy, refered to as von Neumann–Bernays–Gödel (NBG), starts by defining classes and then defines sets as those classes that can be elements of classes.
+It avoids paradoxes by restricting set construction on predicates over elements.
+Since a proper class is not an element, if the definition creates a contradiction for an item $x$ (i.e.
+we have both $x \in S$ and $x \notin S$) than it must be that the element $x$ is a proper class.
+
+\textbf{Axioms}. Here are the axioms of a particular formulation of axiomatic set theory taken from CITE:
+\begin{enumerate}
+	\item Axiom of extent. If two classes have the same elements, then they are identical.
+	\item Axiom of class construction.
+Let $P(x)$ be a predicate expressed in terms of the symbols $\in, \OR, \AND, \NOT, \Rightarrow, \exists, \forall$, brackets, variables $x, y, z, A, B, ...$ Then there exists a class $C$ that consists of all the elements x which satisfy $P(x)$.
+	\item Every subclass of a set is a set.
+	\item $\emptyset$ is a set.
+	\item If $a$ and $b$ are sets, then $\{a, b\}$ is a set.
+	\item If $\mathcal{A}$ is a set of sets, then, $\bigcup\limits_{A \in \mathcal{A}} A$ is a set.
+	\item If $A$ is a set, the power set $\mathcal{P}(A)$ is a set.
+	\item Axiom of foundation. If $A \neq \emptyset$ is a set, there exists an element $a \in A$ such that $a \cap A = \emptyset$.
+	\item Axiom of replacement. If $A$ is a set and $f : A \to B$ is a surjective function, then $B$ is a set.
+	\item Axiom of choice. Every set has a choice function. That is, given a set $A$, there exists a function $f : \mathcal{P}(A) \setminus \{ \emptyset \} \to A$ such that $f(B) \in B$.
+	\item Axiom of infinity. There exists a successor set (as defined in \ref{defn_successor_set}).
+\end{enumerate}
+
+Cardinal numbers
+
+\begin{prop}[Well-ordering theorem]
+	Every set can be well-ordered.
+\end{prop}
+
+Define ordering by existence of injection.
+
+def equipotent set (bijective)
+
+prop no surjective funtion from a to P(A)
+
+talk about contiunuum hypo as example of work in set theory
+
+\begin{remark}
+	Axiom 1 serves to define class equality.
+Axiom 2 defines how to define classes based on predicates.
+Axioms 3-7 tells us that some classes, constructable by axiom 2, are sets.
+Axiom 8 guarantees that we cannot have an infinite sequences $\{ A_i \}_{i=1}^\infty$ such that $A_{i+1} \in A_i$.
+In particular, a set cannot be an element of itself.
+Axiom 9 tells us that any set that is ``smaller'' than a set (i.e.
+we can map another set onto it) is a set.
+Axiom 10 guarantees that we can always pick an element from a set, though it does not tell us how.
+Axiom 11 tells us that there are sets with infinite elements.
+\end{remark}
 

### 2021-05-14 Overviews/Process entropy/ProcessEntropy.tex
ProcEntropy: first pass done

@@
 
-\subsection{Future: States and state spaces}
+\subsection{System, states and state spaces}
 
-In other words, the state space of the system (i.e.
-the set of all possible complete descriptions of the system and only about that system, independently of the process at hand) is something we need to construct.
-It is an endpoint, not a starting point.
-This construction comes with a set of assumptions about the system which are at the foundations of thermodynamics specifically and physics more in general, and they need to be made explicit.
-Best descriptions of a system at a time, instead, do not require any additional assumptions and therefore are more general concepts.
+\textbf{Systems}.
+Lastly we need to answer the following question: how do we go from a set of descriptions to the definition of a system?
+Under what conditions do the snapshot actually represent states?
+And why do we do it?
+The general idea is that, conceptually, we want to study the same object under different processes.
+That is, we want to say that the same object $X$ can be subjected to different process at different intervals of time, and we are able to switch from one to the other.
+Note that this is not simply something we would like to do: it is a basic requirement to do science.
+Process $A$, in fact, can be the reliable preparation of a sample, process $B$ can be some interaction we want to characterize, and process $C$ can be the final measurement.
+What must happen, then, is that the description of the object at each junction is independent from the choice of the previous/future processes and from other objects.
+In other words, the description at the junction is about and only about the system at that time.
+This does not mean that it is the \emph{full} description.
+It must be, then, that we have at our disposal a process under which the object is independent from the rest.
+We removed all correlations in time and with other systems.
+We call an object for which this is possible a \textbf{system}.
 
-\textbf{States and independence}.
-It should be noted that the same description for the same system in different conditions (i.e.
-in different processes or at different times in the same process) will not in general correspond to the same count of evolutions.
+\textbf{Independence as requirement}.
+Therefore \textbf{independence is a requirement for the definition of a system}: we can talk about a system only insofar we have at our disposal a process under which the system is independent for the rest, such that the description of the system is truly a description of the system, with no correlations.
+We can speak of a chair as its own system because we have circumstances in which we can interact with the chair in a way that does not affect the other chairs or the tables.
+The definition of a system, then, is conditional to the existence of suitable processes.
+
+What still needs to be clarified is the difference between the independence from other systems of the system coming in or of the process itself.
+Suppose we have system $X$ and $Y$ whose state is identified by a single continuous variable.
+If there is no correlation, all state combinations are possible, so the joint distribution is uniform over the whole plain.
+Now suppose they are correlated so that that $x=-y$.
+The joint distribution is uniform on the diagonal.
+If we have a process that acts and depends only on $X$, the two cases are indistinguishable since the marginal of $x$ is uniformly distributed in both cases.
+The density of evolutions, however, is different, so there is a renormalization.
+But for the rest, the two cases are indistinguishable.
+So independence may not technically always map to factorization of the measure, but it may need to be indistinguishable from factorization.
+
+\textbf{Independence leads to maximization of entropy}.
+Note that the presence of correlations between the system and the environment can only remove possible evolutions, not add them.
+Therefore independence is the case where the evolution count is maximal.
+In other word, requiring independence is equivalent to requiring maximization of entropy.
+Again note that this maximization of entropy is not due to some property of nature: it is due to the requirement of wanting to define a system, an objects for which we can mix and match processes.
+Of all processes nature presents us, we select those that decouple the world into separate systems, those that maximize entropy.
+
+\textbf{Relaxation processes and equilibria as independence}.
+If we are to study systems, then, we must have at our disposal a process that takes an object and removes all possible correlations, rendering it independent.
+Moreover, this process must keep the system independent if it was already so.
+Therefore \textbf{independent descriptions must be equilibria} of such process.
+In general, the definition of the system requires the existence of such processes, which we call \textbf{relaxation} processes.
+States will be invariant, symmetries, of such processes.
+
+\textbf{System definition as object-environment relationship}.
+Whether or not a particular set of attributes of an object can be seen as an independent system is not an a priori feature of the object, but rather a relationship between object and environment.
+This is not just because of independence, but also for the existence of the system itself.
+We can talk about a ball, with its position and velocity, on the surface of the earth at 20 Celsius.
+On the surface of the sun, we cannot talk about a ball.
+Therefore there must be a tight relationship between states, the processes defined on the system and the object-environment boundary.
+The process cannot alter the boundary, the state space must satisfy the constraints imposed by the boundary, the state space must be, as a whole, a symmetry of the processes.
+
+\textbf{System composition and its algebra}.
+As we may want to describe systems, subsystems and their relationships, it useful to note that \textbf{the set of all possible systems is a Boolean algebra}.
+The ordering of the lattice is given by system inclusion.
+We say $X \in Y$ if $X$ is a subsystem of $Y$.
+We can convince ourselves that this is a partial order (i.e.
+$X \in X$, if $X \in Y$ and $Y \in X$ then $X=Y$ and if $X \in Y$ and $Y \in Z$ then $X \in Z$).
+The join $X \vee Y$ is therefore the smallest system that contains both and represents system composition.
+The meet $X \wedge Y$ is the biggest system contained by both and represent the common subsystem.
+The $1$ represents the system that contains everything.
+The $0$ represent the null system, the one that contains nothing.
+The complement $\neg X$ represents everything except for the system.
+
+\textbf{Carving up the world}.
+Philosophers talk about a description of reality that is joint-carving meaning that reality already contains divisions, prior to our investigation.
+This is an epistemological version of it: nature gives us access to processes for which some systems are independent from the rest.
+It is the accessibility of these processes that allows us to define systems, without them we cannot divide nature into parts.
+That is, if an object is really made of parts but the process driving their evolution is such that the correlations can never be eliminated or controlled, the parts cannot be defined as independent systems.
+
+\textbf{States and state spaces}.
+A \textbf{state is a snapshot for which the system is independent}.
+That is, it is a description of the system and only of the system.
+A \textbf{state space} $X$ is the set of all possible states of the system.
+
+\textbf{States as junction points}.
+States will then act as the juncture point between processes.
+If we switch from process $A$ to process $B$ at time $t_0$, given the requirement of independence, the description of the system at that time must be a state.
+If that's the case, both processes can be characterized independently from each other.
+
+\textbf{States as equilibria of faster/smaller scale processes}.
+As noted above, the subdomain must choose both a physical scale and a time scale.
+The requirement of independence will mean that the description within that scale is not relevant to the processes at hand.
+This cutoff is, in effect, part of the boundary of the system: we are still defining what descriptions, which degrees of freedom, are under consideration and which are not.
+Note that this is consistency of the idea of relaxation process, which must therefore not only remove the correlations from other objects but also from internal degrees of freedom.
+
+\textbf{States as bundles of evolutions over short scale}.
+In this light, is it better to understand states not as points at an instant of time, but rather as bundles of evolutions over short scale that we ``connected'' throughout the process.
+
+\textbf{State entropy}.
+When we join two processes, we need to reconcile the evolution count at the junction.
+This means that the evolution count for states must be well defined and unique if we want to mix and match antecedent and subsequent processes.
+Note that independence is precisely the condition needed for the count of evolutions to factorize and process entropy to becomes additive.
+In this case, we can define a \textbf{state entropy} $S(x)$ which is the process entropy in condition of system independence.
+This means that \textbf{all state spaces must be endowed with a structure that define the state entropy}.
+
+\textbf{The entropic nature of geometric structures of state spaces}.
+The claim is that all geometric structures that state spaces possess in different theories are precisely structures that keep track of the state entropy.
+This is obvious in thermodynamics, where the equation of state in entropric form precisely map state entropy.
+In classical mechanics, phase space is characterized by a symplectic structure, whose volume is used in statistical mechanics to compute the entropy.
+Moreover, orthogonality in the symplectic structure characterize the independence of degrees of freedom.
+In quantum mechanics, the inner product is connected to probabilities that, as we saw, are simply ratios of counts of evolutions.
+Moreover, those same probabilities are used in statistical mechanics for the computation of the entropy.
+Also note that these structure do not provide further characterization of the state space beyond the ability to compute the evolution count and the entropy.
+They provide entropic and only entropic information.
+
+\textbf{Zero state entropy and the null state}.
+As we saw before, entropy is a relative concept and therefore needs a unit to be defined.
+What state should act as a reference and be assigned zero entropy?
+There is a natural choice.
+Compositing with the null system $X \vee 0 = X$ returns the original system, therefore the entropy must not change.
+Since the entropy is linear under system composition, the null system must have zero entropy.
+We should note that, in principle, all systems should have a null state, which is the state where the system is not there.
+This case is automatically independent and is therefore a state.
+This state should be given zero entropy and therefore can be taken as the unit of our measure for the state space.
+
+\textbf{Null state as the lowest entropy state}.
+It should also be intuitive that this state has to have the lowest entropy of all.
+That is, saying the system is not there will give the finest level of description of the system.
+No other description can be finer (except for the impossibility).
+This means that any other state can give a description of the system that is, at best, as fine as the null state.
+This means that the \textbf{the state entropy can never be negative}.
+This is a more general form of the third law of thermodynamics, in the same way that \ref{entropy_determinism} is a more general form of the second law.
+
+\textbf{State entropy and time dependence}.
+We should stress that the state entropy fixes the relative process entropy between states at equal time in conditions of independence.
+In all other cases, or if these conditions changes throughout the process, the connection between state entropy and process entropy is more complicated: the same description for the same system will not correspond to the same count of evolutions.
 Consider a dynamical system with an attractor, such as a damped harmonic oscillator.
@@
 
-If two systems are independent, however, the evolution of each system does not constrain the evolution of the others.
-In this case, the count of evolution, and therefore the entropy, of a particular system is only a function of the description of the independent system.
-The count of evolution factorizes and the entropy becomes additive.
+\textbf{State space as template}.
+We should think of the state space as a template.
+We have a \textbf{state domain} $\tdomain_{X}$ which contain all statement fragments like \statement{the position of the ball is between 2 and 3 meters} together with a surjective map $\iota : \tdomain_{X} \times T \twoheadrightarrow \tdomain_t$ that adds \statement{at time t} for each possible time.
+The map $\iota$ must be a surjection that preserves the logical structure (i.e.
+logical operations, verifiability, narrowness, compatibility, ...) so that the basic logical relationships determined by the system itself are are valid at all times (i.e.
+\statement{at time t the position of the ball is between 2 and 3 meters} $\narrower$ \statement{at time t the position of the ball is between 0 and 300 meters} for every $t$).
 
-What is crucial to understand here is that \textbf{independence is a requirement for the definition of a system}: we can talk about a system only insofar we have at our disposal a process under which the system is independent for the rest, such that the description of the system is truly a description of the system, with no correlations.
-We can speak of a chair as its own system because we have circumstances in which we can interact with the chair in a way that does not affect the other chairs or the tables.
+Note, though, that $\iota( \cdot, t_0)$ in general is not an isomorphism (i.e.
+a bijection that preserves the logical structure) for two reasons.
+The first is that not all states will be available at every moment in every process.
+As we said before, if the process is dissipative the set of possible states will become smaller.
+In that case, $\iota$ will map all the inaccessible states to $\contradiction$, as they will be impossible.
+Therefore $\iota$ is not in general a bijection.
+The second reason is that the statement fragments in $\tdomain_{X}$ represent only descriptions about the system itself and nothing else.
+In a particular process, however, knowing the state of one system might tell us something about other system as well.
+The presence of correlations/coupling may mean that the precise knowledge of the degrees of freedom of the system may allow us to tell something about the environment or the internal structure (i.e.
+microstate).
+Mathematically, $\iota$ may map to statements that are narrower than the original fragment (e.g.
+we can have $\iota($\statement{position of A is 1 meter}$, t_0) \equiv$ \statement{at time $t_0$ the position of A is 1 meter and the position of B is -1 meter}).
 
-Philosophers talk about description of reality that is joint-carving meaning that reality already contains divisions, prior to our investigation.
-This is an epistemological version of it: nature gives us access to processes for which some systems are independent from the rest.
-It is the accessibility of these processes that allows us to define systems, without them we cannot divide nature into parts.
-That is, if an object is really made of parts but the process driving their evolution is such that the correlations can never be eliminated in a controlled fashion, the parts cannot be defined as independent systems.
+\textbf{State spaces are constructed}.
+Though state spaces are template, conceptually they are not the starting point.
+The state space of the system (i.e.
+the set of all possible complete descriptions of the system and only about that system, independently of the process at hand) is something we need to construct.
+We find processes for which the system decouples, and we elevate those descriptions to independent objects.
+This construction comes with a set of assumptions about the system which are at the foundations of thermodynamics specifically and physics more in general, and they need to be made explicit.
+The snapshots (i.e.
+best descriptions of a system at a time), instead, do not require any additional assumptions and therefore are more general concepts.
 
-\section{Current attempt}
+\textbf{A rich mathematical structure} To sum up, the definition of a system, then, must define several things.
+\begin{itemize}
+	\item A boundary, that tells us where the system ends and the environment begins
+	\item A characterization of the boundary, that defines in what cases the system remains well defined
+	\item A set of relaxation processes, that tells us how the system can be decoupled from the rest
+	\item A set of possible independent descriptions (i.e. states), that can be characterized about the system given the circumstances
+	\item A set of possible evolutions for the system (i.e. the processes) under the given circumstances.
+\end{itemize}
+There is a lot more work to be done to fully characterize this mathematical structure.
 
-The math.
+%\section{Current attempt}
 
-\section{Examples}
+%\section{Examples}
 
-\section{Insightful failures}
+%\section{Insightful failures}
 

### 2021-05-14 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: minor comments for Andrea

@@
 
-From a mathematical perspective, the distributive lattice will be provided in both classical and quantum case by a Borel algebra, which is the smallest $\sigma$-algebra on a set $X$ equipped with a topological structure that contains the topology; the non-distributive lattice will be provided by the set of subgroups ordered by inclusion.
+From a mathematical perspective, the distributive lattice will be provided in both classical and quantum case by a Borel algebra, which is the smallest $\sigma$-algebra on a set $X$ equipped with a topological structure that contains the topology; the non-distributive lattice will be provided by the set of \textbf{GC: closed} subgroups ordered by inclusion.
 We will claim that the non-distributive lattice is a strict subset of the distributive one; this result entails that everything that can be described in quantum logic can also be described in classical logic.
@@
 Hence, measurements of cross sections, the most typical type of measurement in the context of particle physics, are not part of the quantum lattice, and neither are measurements of the probabilities themselves.
-Then, to show that one may apply a distributive lattice to quantum propositions can be seen as an important result in order to extend the set of meaningful logical propositions of quantum mechanics.
-To provide an extended quantum logic lies beyond the introductory scope of the present paper and will be the subject for future research; here we limit ourselves to show that such an extension is possible.
+\textbf{GC: La costruzione non \'e chiara...
+forse: Therefore, showing} that one may apply a distributive lattice to quantum propositions can be seen as an important result in order to extend the set of meaningful logical propositions of quantum mechanics.
+\textbf{GC: Providing} an extended quantum logic lies beyond the introductory scope of the present paper and will be the subject for future research; here we limit ourselves to show that such an extension is possible.
+\textbf{GC: Non ci \'e chiaro cosa voglia dire...
+ma forse \'e solo una frase tattica per dire `non aspettative che risolviamo tutto'?}
 
@@
 
-As anticipated above, in this section our arguments are fundamentally based on the notion of $\sigma$-algebra, which is a significant structure in the context of foundational research, given that physically relevant mathematical branches like measure theory (cf.\ \cite{Cohn:2013}) and probability theory (cf.\ \cite{Grimmett:2001}) are built on top of it.
+As anticipated above, in this section our arguments are fundamentally based on the notion of $\sigma$-algebra, which is a significant structure \textbf{GC: from a foundational perspective (non era chiaro cosa foundational research sia...
+ma se \'e un termine tecnico, ricambia pure)}, given that physically relevant mathematical branches like measure theory (cf.\ \cite{Cohn:2013}) and probability theory (cf.\ \cite{Grimmett:2001}) are built on top of it.
 Let us then review some of its essential features.
@@
 As we have already stated, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.\footnote{The choice of the square root of the density over the density itself is merely a technical choice.} In this case we have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
-In all normed spaces, the topology is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.
+In all normed spaces, the topology \textbf{GC: , and therefore the $\sigma$-algebra,} is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.
 
@@
 This is not a closed sub-space and therefore $U \notin L(H)$.
-\textcolor{red}{As the topology of a Hilbert space is Hausdorff, all singletons are closed and therefore $U \in \Sigma$.
-This means $L(H) \subset \Sigma$.
-}This concludes the proof.
+As the topology of a Hilbert space is Hausdorff, all singletons are closed and therefore $U \in \Sigma$.\footnote{To give more context, the set $U = \{ v \}$ can be written as $U = \{w \; | \; |w - v|^2 = 0\}$ the set of all elements with zero distance from $v$.
+This can be understood as the limit (i.e.
+countable intersection) of a sequence of propositions ``the object is within $\epsilon$ of $v$'' where $\epsilon$ tends to zero.
+\textbf{GC: Ho messo come note, cos\`i la dimostrazione rimane visibilmente piccola.
+Vedi te se metterla nel testo.} } This means $L(H) \subset \Sigma$.
+This concludes the proof.
 
@@
 quantum logic) fails to be a Boolean algebra.
-Stone's representation theorem tells us that any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.
+Stone's representation theorem \textbf{GC: serve citazione qui?
+Si pu\`o fare Davey...
+Oppure il paper originale...
+} tells us that any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.
 intersection, union and complement).
@@
 We want to be clear that we are not claiming that the lattice of subspaces of quantum logic is not of interest, or that it is of equal interest in both classical and quantum mechanics.
-Here we have pointed it out that the argument that quantum mechanics \emph{requires} necessarily a radical departure from classical logic is untenable, given that both theories are built on similar structures.
+Here we have pointed \textbf{GC: tolto "it" } out that the argument that quantum mechanics \emph{requires} necessarily a radical departure from classical logic is untenable, given that both theories are built on similar structures.
 We conclude that such claims should be substantially weakened, in the light of the arguments that we proposed in this section.

### 2021-05-17 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: adding more about ordinals and cardinals

@@
 
-\subsection{Ordinal numbers}
-
-For infinite sets, cardinals and ordinals are different since sets of the same infinite size can be given different orderings.
-
+\section{Transfinite (i.e. infinite) numbers}
 \begin{remark}
@@
 first, second, third) identify a position within a sequence.
-Well-ordered sets extend the notion of sequence to arbitrary levels of infinity.
+Cardinals and ordinals are in one to one correspondence only for finite values.
+\end{remark}
+
+\subsection{Ordinal numbers}
+
+\begin{remark}	
+	To define ordinals, we extend the notion of sequences to arbitrary levels of infinity using well-ordered sets.
 \end{remark}
@@
 \begin{defn}[POS]
-	A \textbf{partially ordered set}, \marginleft{POS: $(A, \leq_A)$} or \textbf{POS}, is a tuple $(A, \leq_A)$ where $A$ is a set and $\leq_A$ is a partial order over $A$.
+	A \textbf{partially ordered set}, \marginleft{POS: $(A, \leq_A)$} or \textbf{POS}, is a set with an associated partial order.
+That is, a tuple $(A, \leq_A)$ where $A$ is a set and $\leq_A$ is a partial order over $A$.
 A \textbf{linearly ordered set} is a POS whose order is linear.
@@
 \begin{defn}[Initial segment]
-	The \textbf{initial segment} \marginleft{Initial segment} of a POS $(A, \leq_A)$ determined by $a \in A$ is the set of all elements less than $a$. That is, $S_a=\{ x \in A \, | \, x < a \}$.
+	The \textbf{initial segment} \marginleft{Initial segment} of a POS $(A, \leq_A)$ determined by $a \in A$ is the set of all elements before $a$. That is, $S_a=\{ x \in A \, | \, x < a \}$.
 \end{defn}
@@
 That is, $a_1 \leq_A a_2$ implies $f(a_1) \leq_B f(a_2)$.
-An \textbf{order isomorphism} is a bijective order homomorphism whose inverse is also an order homomorphism.
+An \textbf{order isomorphism} is a bijective order homomorphism whose inverse is also an order homomorphism (i.e.
+the POSs have identical order).
 That is, $b_1 \leq_B b_2$ implies $f^{-1}(b_1) \leq_A f^{-1}(b_2)$.
@@
 \begin{defn}[Well-order]
-	A \textbf{well-ordered} \marginleft{Well-order} set is a linearly ordered set $(A, \leq_A)$ where each subset has a least element.
+	A \textbf{well-ordered} \marginleft{Well-order} set is a linearly ordered set $(A, \leq_A)$ where each subset has a least (i.e.
+first) element.
 That is, for every $B \subseteq A$, there exists an $a \in B$ such that $a \leq_A b$ for all $b \in B$.
-Let $A$ and $B$ be two well ordered set, we say $A$ \textbf{has lower ordinality than} $B$, noted $A \ordleq B$, if $A$ is order isomorphic to an initial segment of $B$.
-They have \textbf{the same ordinality}, noted $A \ordeq B$, if they are order isomorphic.
+Let $A$ and $B$ be two well ordered set, we say $A$ \textbf{has lower ordinality than} $B$, noted $A \ordleq B$, if $A$ is orderwise identical to the beginning of $B$.
+That is, $A$ is order isomorphic to an initial segment of $B$.
+They have \textbf{the same ordinality}, noted $A \ordeq B$, if they have the same order.
+That is, they are order isomorphic.
 \end{defn}
@@
 \begin{prop}[Ordinality]
-	If $A$ and $B$ are two well ordered sets, then either $A \ordleq B$ or $B \ordleq A$. If $\mathcal{A}$ is a collection of well-ordered sets, then one element has least ordinality.
+	If $A$ and $B$ are two well ordered sets, one of them has lower cardinality than the other.
+That is, either $A \ordleq B$ or $B \ordleq A$.
+If $\mathcal{A}$ is a collection of well-ordered sets, then one element has least ordinality.
 \end{prop}
@@
 \begin{remark}
-	Examples of well-ordered sets:
+	Simple examples of well-ordered sets:
 	\begin{itemize}
 		\item a finite sequence: $(a_1, a_2, ..., a_n)$
-		\item an infinite sequence: $(a_1, a_2, ..., a_n, ...)$
-		\item an infinite sequence with elements afterwards: $(a_1, a_2, ..., a_n, ..., b_1, b_2, ... , b_6)$
+		\item an infinite sequence: $(a_1, a_2, ...)$
+		\item an infinite sequence with elements afterwards: $(a_1, a_2, ..., b_1, b_2, ... , b_6)$
+		\item three infinite sequences, one after the other: $(a_1, a_2, ..., b_1, b_2, ... , c_1, c_2, ...)$
 	\end{itemize}
@@
 
-\begin{remark}
-	Ordinal numbers generalize the construction of von Neumann ordinals and cover all possible positions in well-ordered sets.
-\end{remark}
-
 \begin{defn}[Ordinals]
-	An \textbf{ordinal} \marginleft{Ordinals: \ordinals} is a set $A$ with the following properties for all $a, b, c \in A$:
+	An \textbf{ordinal} \marginleft{Ordinals: \ordinals} is a set $A$ that includes the previous and only the previous ordinals. That is, for all elements $a, b, c \in A$ (i.e. previous ordinals of $A$):
 	\begin{enumerate}
-		\item $a \notin a$
-		\item $a \in b$ implies $b \notin a$
-		\item $a \in b$ and $b \in c$ implies $a \in c$
-		\item $x \in a$ implies $x \in A$
+		\item $a \notin a$ (each previous ordinal is not before itself)
+		\item $a \in b$ implies $b \notin a$ (if a previous ordinal is before another, then the latter is not before the former)
+		\item $a \in b$ and $b \in c$ implies $a \in c$ (previous ordinals are transitive)
+		\item $x \in a$ implies $x \in A$ (an ordinal before a previous ordinal is a previous ordinal)
 	\end{enumerate}
@@
 
-\begin{prop}[Ordinals are well-ordered]
-	Let $a \in \ordinals$ be an ordinal. Then $(a, \in)$ is a well-ordered set.
+\begin{prop}[Natural numbers are ordinals]
+	The natural numbers are ordinals. That is, $\mathbb{N} \subset \ordinals$.
 \end{prop}
 
-\begin{prop}[Ordinality of a well-ordered set]
-	For every \marginleft{$\langle A \rangle$} well-ordered set $A$ there is a unique $\langle A \rangle \in \cardinals$ such that $a \ordeq A$.
+\begin{remark}
+	Ordinal numbers so defined generalize the construction of von Neumann ordinals (i.e. a natural number is a set that includes all previous natural numbers) to all possible well-orders.
+\end{remark}
+
+\begin{prop}[Ordinals are well-ordered]
+	All ordinals are well-ordered sets. That is, for each $A \in \ordinals$, $(A, \leq_A)$ is a well-ordered set where $a \leq b$ if either $a \in b$ or $a = b$.
+\end{prop}
+
+\begin{prop}[Well-orders are ordinals]
+	For every \marginleft{$\langle A \rangle$} well-ordered set $A$ there is a unique ordinal that fully characterizes its order.
+That is, there is a unique $\langle A \rangle \in \cardinals$ such that $a \ordeq A$.
 We say $\langle A \rangle$ is the \textbf{ordinality of $A$}.
-The natural numbers are the initial segment of the ordinals.
-$\omega$ is the ordinal associated with the natural numbers.
 \end{prop}
@@
 	As an example, consider the von Neumann ordinal $3 = (0, 1, 2)$ as a well-ordered set.
-We can see it as a sequence of three element, which means the last is the third.
-Similarly, the ordinality of the set $(a_1, a_2, a_3)$ is $3 = (0, 1, 2)$ as they define well-ordered set with the same ordering.
+Similarly, the ordinality of the well-ordered set $(a_1, a_2, a_3)$ is $3 = (0, 1, 2)$ as share the same ordering.
 The definitions are such that these ideas extend to infinite sets.
@@
 \begin{defn}[Ordinal addition]
-	Let $(A, \leq_A)$ \marginleft{Ordinal addition: $\langle A \rangle + \langle B \rangle$} and $(B, \leq_B)$ be two well ordered sets. Let $(A \sqcup B, \leq)$ be the POS where:
+	The addition \marginleft{Ordinal addition: $\langle A \rangle + \langle B \rangle$} of two ordinals is defined by concatenating the first order with the second.
+Let $(A, \leq_A)$ and $(B, \leq_B)$ be two well ordered sets.
+Let $(A \sqcup B, \leq)$ be the POS where:
 	\begin{itemize}
@@
 $\langle A \rangle + \langle B \rangle \neq \langle B \rangle + \langle A \rangle$) for infinite ordinals.
-For example $(a_1, a_2, ..., a_n, ...) \sqcup (b_1, b_2, b_3) = (a_1, a_2, ..., a_n, ..., b_1, b_2, b_3)$ is not the same as $(b_1, b_2, b_3) \sqcup (a_1, a_2, ..., a_n, ...) = (b_1, b_2, b_3, a_1, a_2, ..., a_n, ...) \simeq (c_1, c_2, ..., c_n, ...)$.
+For example $(a_1, a_2, ...) \sqcup (b_1, b_2, b_3) = (a_1, a_2, ..., b_1, b_2, b_3)$ is not the same as $(b_1, b_2, b_3) \sqcup (a_1, a_2, ...) = (b_1, b_2, b_3, a_1, a_2, ...) \ordeq (c_1, c_2, ...)$.
 Therefore $\omega + 3 \neq 3 + \omega = \omega$.
@@
 \begin{defn}[Ordinal multiplication]
-	Let $(A, \leq_A)$ \marginleft{Ordinal multiplication: $\langle A \rangle \langle B \rangle$} and $(B, \leq_B)$ be two well ordered sets. Let $(A \times B, \leq)$ be the POS where:
+	The multiplication \marginleft{Ordinal multiplication: $\langle A \rangle \langle B \rangle$} of two ordinals is defined by ordering their Cartesian product.
+Let $(A, \leq_A)$  and $(B, \leq_B)$ be two well ordered sets.
+Let $(A \times B, \leq)$ be the POS where:
 	\begin{itemize}
@@
 $\langle A \rangle \langle B \rangle \neq \langle B \rangle \langle A \rangle$) for infinite ordinals.
-For example $(a_1, a_2, ..., a_n, ...) \times (b_1, b_2) = ((a_1, b_1), (a_2, b_1), ..., (a_n, b_1), ..., \\ (a_1, b_2), (a_2, b_2), ..., (a_n, b_2), ...)$ is not the same as $(b_1, b_2) \times (a_1, a_2, ..., a_n, ...) = ((b_1, a_1), (b_2, a_1), (b_1, a_2), (b_2, a_2), ..., (b_1, a_n), (b_2, a_n), ...) \simeq (c_1, c_2, ..., c_n, ...)$.
-\\ Therefore $\omega 2 \neq 2 \omega = \omega$.
+For example $(a_1, a_2, ...) \times (b_1, b_2) = ((a_1, b_1), (a_2, b_1), ..., (a_1, b_2), (a_2, b_2), \\...)$ is not the same as $(b_1, b_2) \times (a_1, a_2, ..., a_n, ...) = ((b_1, a_1), (b_2, a_1), (b_1, a_2), \\ (b_2, a_2), ...) \ordeq (c_1, c_2, ...)$.
+Therefore $\omega 2 \neq 2 \omega = \omega$.
 
 	With addition and multiplication, we can get a sense of how the ordinals work.
-$\omega$ represents a countable sequence.
+$\omega$ represents a countable sequence (i.e.
+$\omega = (0, 1, 2, ...) = (\mathbb{N}, \leq)$).
 $\omega +3$ is a countable sequence followed by three elements.
@@
 \end{remark}
+
+\subsection{Cardinal numbers}
 
@@
 \begin{prop}[Cardinal comparability of sets]
-	Given two sets $A$ and $B$, either $A \crdleq B$ or $B \crdleq A$ one must have lower cardinality than the other (i.e.
-there exists an injective function from one two the other).
-Additionally, $A \crdeq B$ if and only if $A \crdleq B$ and $B \crdleq A$ (i.e.
-there exists a bijective function between $A$ and $B$ only if there exists an injective function in both direction).
+	Given two sets $A$ and $B$, one must have lower cardinality than the other.
+That is, either $A \crdleq B$ or $B \crdleq A$.
+Additionally, there exists a bijective function between $A$ and $B$ only if there exists an injective function in both direction.
+That is, $A \crdeq B$ if and only if $A \crdleq B$ and $B \crdleq A$.
 \end{prop}
@@
 That is, $a \ordleq b$ for all $b \in \ordinals$ such that $a \crdeq b$.
-The set of all cardinals is noted $\cardinals \subset \ordinals$.
+The set of all cardinals is noted $\cardinals$ and, by definition $\cardinals \subset \ordinals$.
 \end{defn}
@@
 \begin{prop}[Cardinality of a set]
-	For every \marginleft{$|A|$} set $A$ there is a unique $|A| \in \cardinals$ such that $a \crdeq A$. We say $|A|$ is the \textbf{cardinality of $A$}.
+	For every \marginleft{$|A|$} set $A$ there is a unique cardinal that identifies the cardinality of $A$.
+That is, there exists a unique $|A| \in \cardinals$ such that $a \crdeq A$.
+We say $|A|$ is the \textbf{cardinality of $A$}.
 \end{prop}
 
+\begin{prop}[Cantor's theorem]
+	A set has strictly lower cardinality than its power set. That is, $|A| < |\mathcal{P}(A)|$. If $a = |A|$ then we note $|\mathcal{P}(A)| = 2^a$
+\end{prop}
+
+\begin{prop}[Aleph numbers]
+	The cardinality \marginleft{Aleph \\numbers: $\aleph_a$} of infinite sets are well-ordered. They are noted $\aleph_a$ where $a \in \ordinals$.
+\end{prop}
+
+\begin{defn}[Aleph-nought]
+	The smallest infinite cardinal \marginleft{$\aleph_0$} $\aleph_0$ denotes countable infinity.
+\end{defn}
+
+\begin{defn}[Cardinality of continuum]
+	The cardinality \marginleft{Continuum: $2^{\aleph_0}$} of the continuum is the cardinality of the powerset of a countable set $2^{\aleph_0}$.
+\end{defn}
+
 \begin{remark}
+	Sets that have countable cardinality:
+	\begin{itemize}
+		\item natural $\mathbb{N}$, integer $\mathbb{Z}$ and rational $\mathbb{Q}$ numbers 
+		\item ordinals such as $\omega$, $\omega + 3$, $\omega2$, $\omega^\omega$
+	\end{itemize}
+	Sets that have cardinality of the continuum:
+	\begin{itemize}
+		\item powerset of natural, integers and rationals: $\mathcal{P}(\mathbb{N})$, $\mathcal{P}(\mathbb{Z})$, $\mathcal{P}(\mathbb{Q})$
+		\item real $\mathbb{R}$ and complex $\mathbb{C}$ numbers
+		\item real and complex euclidean spaces: $\mathbb{R}^n$, $\mathbb{C}^n$
+		\item continuous function over euclidean spaces: $C(\mathbb{R}^n, \mathbb{R}^m)$
+		\item standard topology and Borel algebra on $\mathbb{R}^n$ (i.e. the set of open sets and Borel set respectively)
+		\item countable sequences of numbers: $\mathbb{N}^\mathbb{N}$, $\mathbb{Z}^\mathbb{N}$, $\mathbb{Q}^\mathbb{N}$, $\mathbb{R}^\mathbb{N}$, $\mathbb{C}^\mathbb{N}$
+	\end{itemize}
+	Sets that have cardinality greater than the continuum:
+	\begin{itemize}
+		\item powerset of the real and complex numbers: $\mathcal{P}(\mathbb{R})$, $\mathcal{P}(\mathbb{C})$
+		\item set of all functions from $\mathbb{R}^n$ to $\mathbb{R}^m$
+	\end{itemize}
+\end{remark}
+
+\begin{defn}[Aleph-one]
+	Let $\Omega$ \marginleft{$\aleph_1$} be the set of all countable ordinal numbers. Its cardinality $|\Omega| = \aleph_1$ is the second aleph number.
+\end{defn}
+
+\begin{remark}
+	 While it can be proven that the cardinality of the continuum $2^{\aleph_0}$, yet which one is undetermined.
+\end{remark}
+
+\begin{axiom}[Continuum hypothesis]
+	There is \marginleft{Continuum hypothesis} not infinite set with cardinality in between that of the continuum and countable infinity. That is: $2^{\aleph_0} = \aleph_1$.
+\end{axiom}
+
+\begin{remark}
+	The continuum hypothesis is proven to be independent on the axioms of set theory and is therefore an additional optional axiom.
 \end{remark}

### 2021-05-18 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: more work on cardinals, reviewed Naive Set Theory

@@
 
-\def\ordless{\prec}
 \def\ordleq{\preceq}
@@
 \def\ordgeq{\succeq}
-\def\ordgtr{\succ}
 
-\def\crdless{\sqsubset}
 \def\crdleq{\hookrightarrow}
@@
 \def\crdgeq{\hookleftarrow}
-\def\crdgtr{\sqsupset}
 
@@
 \end{abstract}
-
-Ordinal: $\ordless \ordleq \ordeq \ordgeq \ordgtr$
-
-Cardinal: $\crdless \crdleq \crdeq \crdgeq \crdgtr$
 
@@
 	Given a set $A$, its \textbf{power set}, \marginleft{Power set:\\ $\mathcal{P}(A)$, $2^{A}$} noted $\mathcal{P}(A)$ or $2^{A}$, is the set of all subsets of $A$.
-That is, $\mathcal{P}(A) = \{B \, | \, B \subset A \}$.
+That is, $\mathcal{P}(A) = \{B \, | \, B \subseteq A \}$.
 \end{defn}
@@
 
+\begin{remark}
+	Ordered pairs and disjoint unions exemplify the effort in set theory to construct \emph{all object} starting from the primitives of set and set inclusion.
+\end{remark}
+
 \subsection{Relations and functions}
@@
 the relation is the set of pairs $(a, b)$ such that $a=b$ or $a\leq b$ respectively).
-This is an example of highly fruitful generalization.
+This exemplifies the effort in set theory to find general structures.
 \end{remark}
@@
 \begin{defn}[Binary relation]
-	Given a set $A$, \marginleft{Binary\\ relation} called \textbf{domain}, and a set $B$, called \textbf{codomain}, a \textbf{binary relation} is a set $R \subseteq A \times B$ of ordered pairs.
+	Given a set $A$, \marginleft{Binary\\ relation: $aRb$} called \textbf{domain}, and a set $B$, called \textbf{codomain}, a \textbf{binary relation} is a set $R \subseteq A \times B$ of ordered pairs.
 We say $a$ is \textbf{$R$-related} to $b$, noted $aRb$ if $(a,b) \in R$.
@@
 \begin{defn}[Converse relationship]
-	Given a relationship  \marginleft{Converse: $R^{-1}$} $R$ between $A$ and $B$, the \textbf{converse} or \textbf{inverse} relationship $R^{-1}$ is the relationship where the order of the pairs is inverted.
+	Given a relationship  \marginleft{Converse: $R^{-1}$} $R$ between $A$ and $B$, the \textbf{converse} or \textbf{inverse} relationship $R^{-1} \subseteq B \times A$ is the relationship where the order of the pairs (i.e.
+domain and codomain) is inverted.
 That is, $R^{-1} = \{ (b, a) \, | \, (a, b) \in R\}$.
@@
 	\begin{description}
-		\item[Functional or right-unique.] An element $a \in A$ is related to at most one element $b \in B$. That is, if $aRb$ and $cRb$ then $a = c$.
-		\item[Injective or left-unique.] An element $b \in B$ is related to at most one element $a \in A$. That is, if $aRb$ and $aRc$ then $b = c$.
+		\item[Functional or right-unique.] An element $a \in A$ is related to at most one element $b \in B$. That is, if $aRb$ and $aRc$ then $b = c$.
+		\item[Injective or left-unique.] An element $b \in B$ is related to at most one element $a \in A$. That is, if $aRb$ and $cRb$ then $a = c$.
 		\item[Serial or left-total.] Every element $a \in A$ is related to at least one element $b \in B$.
@@
 
-\begin{defn}
-	Let $f : A \to B$ be a function.
-This induces a relation $f : \mathcal{P}(A) \to \mathcal{P}(B)$ that associates each subset of $A$ to its image and a relation $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ that associates each subset of $B$ with its reverse image.
+\begin{defn}[Identity function]
+	Given a set $A$, \marginleft{Identity: $\Id_A$} the identity function $\Id_A : A \to A$ is the function such that $\Id_A(a) = a$ for all $a \in A$.
 \end{defn}
 
-\begin{defn}[Level set]
-	A level set \marginleft{Level set} of a function $f : A \to B$ is a set of where the function takes a given value $b$. In other words the reverse image $f^{-1}(\{c\})$ a singleton.
+\begin{prop}[Inverse function]
+	Let $f : A \to B$ \marginleft{Inverse: $f^{-1}$} be a function.
+The corresponding converse relationship $f^{-1}$ is a function if and only if $f$ is bijective.
+In this case $f^{-1} : B \to A$ is also bijective and is called the \textbf{inverse} of $f$.
+\end{prop}
+
+\begin{defn}[Image/preimage functions]
+	Let $f : A \to B$ \marginleft{$f(X), f^{-1}(Y)$} be a function.
+This induces a relation $f : \mathcal{P}(A) \to \mathcal{P}(B)$ that associates each subset of $A$ to its image and a relation $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ that associates each subset of $B$ with its preimage.
 \end{defn}
 
-\begin{prop}
-	The induces relations $f : \mathcal{P}(A) \to \mathcal{P}(B)$ and $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ are functions and satisfy the following properties:
+\begin{prop}[Properties of image/preimage functions]
+	The induced relations $f : \mathcal{P}(A) \to \mathcal{P}(B)$ and $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ are functions and satisfy the following properties:
 	\begin{enumerate}
@@
 \begin{remark}
-	With abuse of notation, $f$ and$f^{-1}$ indicate both the function between elements and sets.
-The image of the intersection is not in general the intersection of the image.
-Note how in both topology and measure theory continuity and measurability are defined based on the reverse images.
+	With abuse of notation, $f$ and $f^{-1}$ indicate both the function between elements and sets.
+Note that $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ always exists even if $f$ is not bijective.
+Also note that the image of the intersection is not in general the intersection of the image, and how maps that preserve set theoretic structures (e.g.
+topologies and $\sigma$-algebras) are defined based on reverse images.
 \end{remark}
 
-\begin{defn}[Identity function]
-	Given a set $A$, \marginleft{Identity: $\Id_A(a)$} the identity function $\Id_A : A \to A$ is the function such that $\Id_A(a) = a$ for all $a \in A$.
+\begin{defn}[Fiber]
+	Given \marginleft{Fiber} a function $f : A \to B$, the A \textbf{fiber of $b \in B$ under $f$} is the subset of $A$ of where the function takes the value $b$.
+That is, the reverse image $f^{-1}(\{b\})$ of the singleton.
+Fibers over real functions are also called \textbf{level sets}.
 \end{defn}
-
-\begin{prop}[Inverse function]
-	Let $f : A \to B$ \marginleft{Inverse: $f^{-1}$} be a function.
-The corresponding converse relationship $f^{-1}$ is a function if and only if $f$ is bijective.
-In this case $f^{-1} : B \to A$ is also bijective and is called the \textbf{inverse} of $f$.
-\end{prop}
 
@@
 Let $R$ be a binary relationship between $A$ and $B$ and $S$ be a binary relationship between $B$ and $C$.
-Then the \textbf{composition of $R$ and $S$}, noted $S \circ R$, is the binary relationship between $A$ and $C$ such that $(a, c) \in S \circ R$ if we can find $b \in B$ such that $aRb$ and $bSc$.
+Then the \textbf{composition of $R$ and $S$} is the binary relationship $S \circ R \subseteq A \times C$ between $A$ and $C$ such that $(a, c) \in S \circ R$ if we can find $b \in B$ such that $aRb$ and $bSc$.
 \end{defn}
@@
 \begin{remark}
-	The notation is due to the cardinality.
-If $|A| = n$ and $|B| = m$, each function is given by choosing an element of $B$ for each element of $A$, that is $m$ choices $n$ times, that is $m^n$.
+	The notation for function sets is due to the cardinality.
+If $A$ has $n$ elements and $|B|$ has $m$ elements, each function is given by choosing an element of $B$ for each element of $A$, that is $m$ choices $n$ times, that is $m^n$.
 The power set is noted $2^A$ because each subset can be identified by a function $U : A \to \{0 , 1\}$, where $f(a)=0$ corresponds to $a \notin U$ and $f(a)=0$ corresponds to $b \in U$.
@@
 
-TODO: add definition of set closed under operation
+\begin{defn}[Set closure under operations]
+	Let $f : A_1 \times A_2 \times ...
+\times A_n \to B$ be an $n$-ary function.
+We say a set \textbf{$A$ is closed under $f$} if whenever the function is applied to elements of $A$ returns an element of $A$.
+That is, $A \subseteq A_i$ for all $1\leq i \leq n$, $A \subset B$ and $f(A, A, ..., A) \subseteq A$.
+\end{defn}
+
+\begin{remark}
+	For example, integers are closed under addition, subtraction and multiplication but are not closed under division.
+Most mathematical structures (groups, topologies, $\sigma$-algebra, vector spaces, ...) define sets closed under some operations.
+\end{remark}
 
@@
 More formally, an indexed family is a tuple $\langle X, I, x \rangle$ where $X$ and $I$ are sets and $x: I \to X$ is a function such that $i \mapsto x_i$.
-A \textbf{sequence} is an family for which the index set is a continuous interval of natural numbers.
+A \textbf{sequence} is a family for which the index set is a continuous interval of natural numbers.
 A \textbf{finite sequence} of $n$ elements is noted $\{x_i\}_{i=1}^{n}$ while an \textbf{infinite sequence} is noted $\{x_i\}_{i=1}^{\infty}$.
@@
 \begin{defn}[Monotone sequence]
-	A \textbf{monotone sequence} \marginleft{Monotone sequence: \\ $A_1 \subseteq A_2 \subseteq ...$\\ $A_1 \supseteq A_2 \supseteq ...$} is a sequence of sets $\{A_i\}_{i=0}^{\infty}$ where each set is a superset (or subset) of the preceding.
+	A \textbf{monotone sequence} \marginleft{Monotone sequence: \\ $A_1 \subseteq A_2 \subseteq ...$\\ $A_1 \supseteq A_2 \supseteq ...$} is a sequence of sets $\{A_i\}_{i=0}^{\infty}$ where each set is a subset (or superset) of the preceding.
 More specifically, we distinguish four cases:
@@
 \begin{remark}
-	We will cover the notions of orders that are necessary for the definition of ordinals and cardinal numbers. Orders will treated more in detail in a Bare Minimum dedicated to order theory.
+	We will cover the notions of orders that are necessary for the definition of ordinals and cardinal numbers. Orders will be treated more in detail in a Bare Minimum dedicated to order theory.
 \end{remark}
@@
 		\item[Transitive.] If $a$ is related to $b$ and $b$ is related to $c$, then $a$ is related to $c$. That is, if $aRb$ and $bRc$ then $aRc$.
-		\item[Total.] Given two distinct elements, one is related to the other.That is, given $a \neq b$, either $aRb$ or $bRa$.
+		\item[Total.] Given two distinct elements, one is related to the other.That is, given $a \neq b$, either $aRb$ or $bRa$. (This has no relation to left/right-total)
 	\end{description}
@@
 The \textbf{equivalence class of $a$ by $\sim$} is the set $[a]_{\sim} = a_{/\sim} = \{b \in A \, | \, b \sim a \}$ of all the elements of $A$ equivalent to $a$.
-The \textbf{quotient set of $A$ by $\sim$} $A_{/\sim} = \{ a_{/\sim} \, | \, a \in A \}$ all of the equivalences classes by $\sim$.
+The \textbf{quotient set of $A$ by $\sim$} $A_{/\sim} = \{ a_{/\sim} \, | \, a \in A \}$  is the set of all of the equivalences classes by $\sim$.
 \end{defn}
 
-\begin{prop}
+\begin{prop}[Quotient sets are partitions]
 	Every quotient set is a partition, and every partition is a quotient set.
@@
 
-%\begin{prop}[Set inclusion as ordering]
-%	Set inclusion \marginleft{$\subseteq$ and $\supseteq$ are partial orders} (either $\subseteq$ or $\supseteq$) is a partial order.
-Strict set inclusion (either $\subset$ or $\supset$) is a strict partial order.
-%\end{prop}
+\begin{remark}
+	All the above concepts are of general use in all areas of mathematics. Note that they are all constructed on two primitives: the notion of sets and of set inclusion.
+\end{remark}
 
-\subsection{Numeric constructions}
+\section{Numeric constructions}
 \begin{remark}
@@
 	\end{itemize}
+
+	We should note that all physically meaningful sets have at most the cardinality of the continuum.
+This is not a coincidence.
+In science we are interested in object that can be experimentally identified.
+Every experimental test can acquire finite information in finite time and, assuming arbitrarily long time, we can run at most countably many tests.
+Countable sequences of numbers have the cardinality of the continuum, which is therefore the best we can distinguish.
+Even if nature allows for wider variety, it would be merged into equivalence classes of objects that are experimentally indistinguishable.
+Much of issues of higher cardinals can therefore be safely set aside by those working on the Assumptions of Physics.
 \end{remark}
@@
 
-\begin{axiom}[Continuum hypothesis]
+\textbf{Continuum hypothesis}
 	There is \marginleft{Continuum hypothesis} not infinite set with cardinality in between that of the continuum and countable infinity. That is: $2^{\aleph_0} = \aleph_1$.
-\end{axiom}
 
@@
 	The continuum hypothesis is proven to be independent on the axioms of set theory and is therefore an additional optional axiom.
+As the axiom deals with sets of cardinality lower than the continuum, this aspect of set theory may be interesting for those working on the Assumptions of Physics.
+It could be, in principle, that the distinguishable configurations in space-time have a cardinality that is neither countable nor that of the continuum.
 \end{remark}
@@
 
-\textbf{Axioms}. Here are the axioms of a particular formulation of axiomatic set theory taken from CITE:
+\textbf{Axioms of set theory}. Here are the axioms of a particular formulation of axiomatic set theory taken from CITE:
 \begin{enumerate}
@@
 
-Cardinal numbers
-
-\begin{prop}[Well-ordering theorem]
-	Every set can be well-ordered.
-\end{prop}
-
-Define ordering by existence of injection.
-
-def equipotent set (bijective)
-
-prop no surjective funtion from a to P(A)
-
-talk about contiunuum hypo as example of work in set theory
-
 \begin{remark}
@@
 Axiom 11 tells us that there are sets with infinite elements.
+Once the natural numbers are assumed to exist, higher cardinality of infinity can be constructed through, for example, the power set.
+
+	A lot of work in set theory has been and is to clarify the relationship between the axioms, in particular which are independent from the others.
+The axiom of choice has been controversial for some time and therefore has been one of the most studied.
 \end{remark}
 
-Ugly elements of set theory: all mathematical objects are sets. Numbers are set.
+\begin{prop}
+	The following statements are equivalent:
+	\begin{itemize}
+		\item Axiom of choice: every set has a choice function
+		\item Well-ordering theorem: every set can be well ordered
+		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other
+		\item Zorn's lemma: every non-empty POS for which every chain (i.e. totally ordered subset) has an upper bound contains at least one maximal element
+		\item Every vector space has a basis
+	\end{itemize}
+\end{prop}
+
+\begin{remark}
+	It should be stressed that there are different approaches.
+Constructive set theory, as all constructive mathematics, rejects ``nonconstructive proofs'' in which the existence of an object is proved with no explicit algorithm as to how to constructive.
+This approach will, for example, exclude the axiom of choice.
+For those working on the Assumptions of Physics, then, it is important to understand that the foundations of math are not absolute.
+Therefore we have to understand the consequences of those choice and understand whether they are suitable for the foundations of physics.
+\end{remark}
 

### 2021-05-19 Overviews/Informational granularity/InformationGranularity.tex
InfGran: version pushed on arxiv

@@
 Then we can write $\stmt_1 \OR \stmt[v]_1 \eqgran \stmt_2 \OR \stmt[v]_2$ and $\stmt_2 \OR \stmt[w]_2 \eqgran \stmt_3 \OR \stmt[w]_3$.
-If the infinitesimal where incompatible we would have $\stmt_1 \OR \stmt[v]_1 \OR \stmt[w]_2 \eqgran \stmt_2 \OR \stmt[v]_2 \OR \stmt[w]_2 \eqgran \stmt_3 \OR \stmt[w]_3 \OR \stmt[v]_2$.
+If the infinitesimals were incompatible we would have $\stmt_1 \OR \stmt[v]_1 \OR \stmt[w]_2 \eqgran \stmt_2 \OR \stmt[v]_2 \OR \stmt[w]_2 \eqgran \stmt_3 \OR \stmt[w]_3 \OR \stmt[v]_2$.
 If the infinitesimals were comparable, by proposition \ref{infinitesimal_disjunction} we would have that $\stmt[v]_1 \OR \stmt[w]_2$ and $\stmt[w]_3 \OR \stmt[v]_2$ are infinitesimals and therefore $\stmt_1 \equ \stmt_3$.
@@
 It may be that totality is not needed for proposition \ref{infinitesimal_disjunction}.
-It may be that we need to amend the definition of infinitesimals, such that the disjunction of two infinitesimals is an infinitesimals.
-We could also force the transitivity, even if the disjunction of two infinitesimals is not an infinitesimals.
-This would be like saying that, yes, in principles the assertions correspond to different granularity levels, but we are not to distinguish between them because the different is effectively zero.
+It may be that we need to amend the definition of infinitesimals, such that the disjunction of two infinitesimals is an infinitesimal.
+We could also force the transitivity, even if the disjunction of two infinitesimals is not an infinitesimal.
+This would be like saying that, yes, in principle the assertions correspond to different granularity levels, but we are not to distinguish between them because the difference is effectively zero.
 Understanding which options are best requires more work.
@@
 \begin{desid}
-	Let $\stmt_1, \stmt_2 \in \tdomain$ such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$. Then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
+	Let $\stmt_1, \stmt_2 \in \tdomain$ be such that $\stmt_1 \equ \stmt_2$ and $\stmt_1 \finer \stmt_2$. Then for all $\stmt_1 \finer \stmt \finer \stmt_2$ we have $\stmt_1 \equ \stmt \equ \stmt_2$.
 \end{desid}
@@
 \begin{desid}
-	Let $\stmt_1,\stmt,\stmt[v]_1\in\tdomain$ such that $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$, and suppose $\stmt[v]_1$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt\equ\stmt_1$. 
+	Let $\stmt_1,\stmt,\stmt[v]_1\in\tdomain$ be such that $\stmt_1\finer\stmt\finer\stmt_1\OR\stmt[v]_1$, and suppose $\stmt[v]_1$ is infinitesimal with respect to $\stmt[u]$. Then $\stmt\equ\stmt_1$. 
 \end{desid}
@@
 
-Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotone?
+Question: is the map $\tdomain / \eqgran \to \tdomain / \equ$ defined by collapsing equivalence classes well-defined and monotonic?
 
@@
 
-The established results in the literature of comparable probability require a specific axiom, called monotone continuity, that essentially make sure that sequences converge.
+The established results in the literature of comparable probability require a specific axiom, called monotone continuity, that essentially makes sure that sequences converge.
 This is necessary because their goal is to recover a probability measure that is countably additive.
@@
 See examples in \ref{lack_of_monotone_continuity}.
-In terms of measures, the issue is that we may have that the disjoint disjuntion of countably many infinitesimals is not necessarily an infinitesimal, which means countably meany sets of measure zero would not sum up to a set of measure zero.
+In terms of measures, the issue is that we may have that the disjoint disjunction of countably many infinitesimals is not necessarily an infinitesimal, which means countably many sets of measure zero would not sum up to a set of measure zero.
 
@@
 The paper \cite{villegas} by Villegas gives us some conditions under which we can recover a probability measure given a preorder.
-Here we state the conditions required in terms of $\lequ$ and $\tdomain_{\stmt[u]}$, which we take it to be the set of statements narrower than $\stmt[u]$.
+Here we state the conditions required in terms of $\lequ$ and $\tdomain_{\stmt[u]}$, which we take to be the set of statements narrower than $\stmt[u]$.
 We start from the axioms of pre-order:
@@
 Then from $\stmt_1 \lequ \stmt[t]_1$ and $\stmt_2 \lequ \stmt[t]_2$ if follows that $\stmt_1\OR \stmt_2 \lequ \stmt[t]_1\OR \stmt[t]_2$.
-Moreover, if $\lequ$ is replace by $\ltu$ in one of the first two equality, then the last one holds with $\ltu$.
+Moreover, if $\lequ$ is replaced by $\ltu$ in one of the first two inequalities, then the last one holds with $\ltu$.
 
@@
 Therefore $D_{i,j} \eqgran D_{i,l}$ for any $j$ and $l$.
-We have $1 = \mu_{\mathbb{N}}(\mathbb{N}) = \mu_{\mathbb{N}}(\bigOR_{j=1}^i D_{i,j}) = \sum_{j=1}^i \mu_{\mathbb{N}}( D_{i,j}) = \sum_{j=1}^i \mu_{\mathbb{N}}( D_{i,0}) = i \mu_{\mathbb{N}}( D_{i,0}) $.
+We have $1 = \mu_{\mathbb{N}}(\mathbb{N}) = \mu_{\mathbb{N}}(\bigOR_{j=0}^{i-1} D_{i,j}) = \sum_{j=0}^{i-1} \mu_{\mathbb{N}}( D_{i,j}) = \sum_{j=0}^{i-1} \mu_{\mathbb{N}}( D_{i,0}) = i \mu_{\mathbb{N}}( D_{i,0}) $.
 Therefore $\mu_{\mathbb{N}}( D_{i,j}) = 1/i$ for any $i$ and $j$.
@@
 Consider the reals with the standard measure.
-A set of finitely many points points has measure zero.
+A set of finitely many points has measure zero.
 The rationals are countable and have measure zero.
-The Cantor set is uncountable and have measure zero.
+The Cantor set is uncountable and has measure zero.
 Therefore sets of real numbers with all possible cardinality can have measure zero.
@@
 If we are treating $\mathbb{N}$ as a uniform domain, we should have $\stmt_n\eqgran\stmt[t]_n$ for all $n$.
-However, the limit $\bigOR_n\stmt_n = \mathbb{N}$, while $\bigOR_n\stmt[t]_n = \{2,4,6,\ldots\} = 2\mathbb{N} \subset \mathbb{N}$.
+However, the limit $\bigOR_n\stmt_n = \mathbb{N}$, while $\bigOR_n\stmt[t]_n = \{2,4,6,\ldots\} \subset \mathbb{N}$.
 These are not equigranular limits despite the sequences being equigranular.
@@
 For example, if we have a countable set of possibilities (e.g.
-the natural numbers), a saturated chain in narrowness can start from the empty set, start by adding possibilities one be one until we end with the full set.
+the natural numbers), a saturated chain in narrowness can start from the empty set, start by adding possibilities one by one until we end with the full set.
 The set with all possibilities except one, for example, cannot be added to the chain, but still represents a level of description that is in between all the finite sets and the full infinite set.
@@
 A theoretical domain $\tdomain$ is said to have the \textbf{saturated chains property} (SCP) if given any chain $S\subseteq\tdomain$ totally ordered by narrowness ($\narrower$), the following holds.
-There exists a maximum chain $S'\supseteq S$, and a saturated chain $C$ in the set of equigranularity equivalence classes, such that for all statements $s \in S'$, there exists an equivalence class $\overline{c}\in C$ with $s\in \overline{c}$.
+There exists a saturated chain $S'\supseteq S$, and a saturated chain $C$ in the set of equigranularity equivalence classes, such that for all statements $s \in S'$, there exists an equivalence class $\overline{c}\in C$ with $s\in \overline{c}$.
 \end{defn}
@@
 By the SCP, these may be extended to saturated chains in narrowness, with corresponding saturated chains in fineness equivalence classes.
-Then since possibilities are successors of the impossibility, the fineness equivalence classes represent by $p$ and $q$ must both be an immediate successor of the equivalence class representing the impossibility.
+Then since possibilities are successors of the impossibility, the fineness equivalence classes represented by $p$ and $q$ must both be an immediate successor of the equivalence class representing the impossibility.
 It follows that $p\finer q$, and similarly $q\finer p$, and so $p\eqgran q$.
@@
 \begin{prop}
-Let $\tdomain$ be a theoretical domain with the saturated chains property. Then they must have finitely many possibilities.
+Let $\tdomain$ be a theoretical domain with the saturated chains property. Then it must have finitely many possibilities.
 \end{prop}
@@
 Note that $A$ cannot contain all previous $x_j$ in the sequence, since $\hat{\stmt} \notin C$.
-Let $x_j$ the possibility with the highest $j \leq i$ that is not in $A$.
+Let $x_j$ be the possibility with the highest $j \leq i$ that is not in $A$.
 We have $\hat{\stmt} \nnarrower \stmt_{j}$ and $\stmt_{i} \nnarrower \hat{\stmt}$ since each is compatible with a possibility incompatible with the other.
-Since we cannot add other statement, $C$ is saturated.
+Since we cannot add other statements, $C$ is saturated.
 
@@
 Let $x \in X'$ and consider $\stmt \AND \NOT x$.
-This is finer than $\stmt$ and is coarser than all other statements statements in $C$ since they are all compatible with finitely many possibilities.
+This is finer than $\stmt$ and is coarser than all other statements in $C$ since they are all compatible with finitely many possibilities.
 Therefore SCP does not hold.

### 2021-05-19 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: draft finished

@@
 
+\def\ordless{\prec}
 \def\ordleq{\preceq}
@@
 \begin{defn}[Disjoint union]
-	Let $A$ and $B$ be two sets. The disjoint union $A \sqcup B$ is the union where the elements of the sets are always treated as distinct. That is, $A \sqcup B = \{A\} \times A \cup \{B\} \times B$.
+	Let $A$ and $B$ be two sets.
+The disjoint union $A \sqcup B$ is the union where the elements of the sets are always treated as distinct.
+That is, given $A \sqcup B = \{0\} \times A \cup \{1\} \times B$.
 \end{defn}
@@
 
+\section{Axiomatic set theory}
+
+Axiomatic set theory was introduced to eliminate some technical problems that would make naive set theory inconsistent.
+There is no single approach and there is no single set of axioms.
+For example, constructive set theory, as all constructive mathematics, rejects ``nonconstructive proofs'' in which the existence of an object is proved with no explicit algorithm as to how to construct it.
+This will lead to different choices than the ones outlined here.
+Much of the work on the foundations of mathematics in set theory is to clarify the relationship between the axioms: which ones are required to get certain results, which ones are independent and which ones are inconsistent.
+
+This is an important insight for those working on the Assumptions of Physics: the foundations of mathematics are not absolute.
+We have to be conscious of the consequences of those choice and understand whether they are suitable for the foundations of physics.
+
+\subsection{Paradoxes}
+
+While they are generally called paradoxes, the following are really contradictions and that is why axiomatic set theory needs to avoid them.
+The first type of paradoxes to naive set theory arise from the use of natural language and the meaning of the words, typically when this is self-referential.
+
+\textbf{Berry's paradox}.
+Consider the expression ``the smallest positive integer not definable in under sixty letters.'' Given that there are only twenty-six letters in the English language, there are finitely many phrases under sixty letters and therefore only finitely many integers can be defined in under sixty letters.
+Therefore there are integers not definable under sixty letters and, since the integers are ordered, there will be a least one.
+However, the above expression contains fifty-seven letters.
+We have a contradiction.
+
+These type of paradoxes are solved by restricting to symbolic expressions of a formal language, such as predicate logic.
+The consequence is that, at the foundational level, mathematics does not keep track of what the symbols mean.
+
+The first type of paradoxes are purely logical, so they can be constructed in the formal system.
+
+\textbf{Russell's paradox}.
+Let $R$ be the set of all sets that contain themselves.
+That is, $R=\{ x \, | \, x \notin x\}$.
+Suppose $R \notin R$ then, by definition, $R \in R$.
+Suppose $R \in R$ then, by definition, $R \notin R$.
+We have a contradiction.
+
+These are solved by carefully specifying the rules that are allowed for the creation of sets, by picking the ``right'' axioms. Part of the solution revolves around the following concepts.
+
+%https://en.wikipedia.org/wiki/Zermelo%E2%80%93Fraenkel_set_theory
+
+%https://en.wikipedia.org/wiki/Von_Neumann–Bernays–Gödel_set_theory
+
+\subsection{Classes, sets and elements}
+
+A \textbf{class} \marginright{TODO: diagram with Classes, proper classes and sets.}  is any collection, therefore \textbf{sets are classes}.
+Some classes are not sets.
+These are called \textbf{proper classes} and are exactly the collections that cannot be elements of a class.
+Therefore, sets are classes that can be contained by other classes, while proper classes cannot.
+The difference between sets and proper classes is not in what they contain but in whether they can be contained.
+Proper classes are ``too large'' to be sets in the following sense: given a class $A$, if we can find a set $B$ that contains all elements of $A$, then $A$ will be a set as well.
+Therefore proper classes can never be subclasses of sets, they are bigger than all sets.
+The class of all sets, for example, is a proper class.
+So is the class of all cardinal numbers.
+
+Note that axiomatic set theory does not distinguish between objects that can or cannot contain other objects.
+All elements are classes, all element can, in principle, contain something.
+In fact, since equality of two objects coincides on whether they have the same elements, two objects must contain something different to be different.
+Numbers will be constructed as sets starting from the empty set.
+As a consequence, the proposition $3 \in 4$ is valid, and happens to be true if $3$ and $4$ are naturals, though it is not true if $3$ and $4$ are reals.
+From the perspective of the Assumptions of Physics, this is not great as the framework allows propositions that are not physically meaningful.
+Radical different approaches to the foundations of mathematics (e.g.
+through category theory) do not exhibit this particular problem, though they exhibit others (e.g.
+identity of an object can be defined only up to its relationships with other objects).
+
+\subsection{Axiomatization}
+
+One strategy for the axiomatization of set theory, refered to as Zermelo–Fraenkel (ZFC), only defines sets.
+It avoids paradoxes by restricting set construction to subsets only.
+Proper classes are not objects that live in the framework.
+
+Another strategy, refered to as von Neumann–Bernays–Gödel (NBG), starts by defining classes and then defines sets as those classes that can be elements of classes.
+It avoids paradoxes by restricting set construction on predicates over sets.
+Since a proper class cannot be an element, if the definition creates a contradiction for an item $x$ (i.e.
+we have both $x \in S$ and $x \notin S$) than it must be that the element $x$ is a proper class.
+
+The two approaches are proved to be equivalent.
+
+\textbf{Axioms of set theory}. Here are the axioms of a particular formulation of axiomatic set theory taken from CITE that follows NBG:
+\begin{enumerate}
+	\item Axiom of extent. If two classes have the same elements, then they are identical.
+	\item Axiom of class construction.
+Let $P(x)$ be a predicate expressed in terms of the symbols $\in, \OR, \AND, \NOT, \Rightarrow, \exists, \forall$, brackets, variables $x, y, z, A, B, ...$ Then there exists a class $C$ that consists of all the elements (i.e.
+sets) x which satisfy $P(x)$.
+	\item Every subclass of a set is a set.
+	\item $\emptyset$ is a set.
+	\item If $a$ and $b$ are sets, then $\{a, b\}$ is a set.
+	\item If $\mathcal{A}$ is a set of sets, then, $\bigcup\limits_{A \in \mathcal{A}} A$ is a set.
+	\item If $A$ is a set, the power set $\mathcal{P}(A)$ is a set.
+	\item Axiom of foundation. If $A \neq \emptyset$ is a set, there exists an element $a \in A$ such that $a \cap A = \emptyset$.
+	\item Axiom of replacement. If $A$ is a set and $f : A \to B$ is a surjective function, then $B$ is a set.
+	\item Axiom of choice. Every set has a choice function. That is, given a set $A$, there exists a function $f : \mathcal{P}(A) \setminus \{ \emptyset \} \to A$ such that $f(B) \in B$.
+	\item Axiom of infinity. There exists a successor set (as defined in \ref{defn_successor_set}).
+\end{enumerate}
+
+\begin{remark}
+	Axiom 1 serves to define class equality.
+Axiom 2 defines how to define classes based on predicates.
+Axioms 3-7 tells us that some classes, constructable by axiom 2, are sets.
+Axiom 8 guarantees that we cannot have an infinite sequences $\{ A_i \}_{i=1}^\infty$ such that $A_{i+1} \in A_i$.
+In particular, a set cannot be an element of itself.
+Axiom 9 tells us that any set that is ``smaller'' than a set (i.e.
+we can map another set onto it) is a set.
+Axiom 10 guarantees that we can always pick an element from a set, though it does not tell us how.
+Axiom 11 tells us that there are sets with infinite elements.
+Once the natural numbers are assumed to exist, higher cardinality of infinity can be constructed through, for example, the power set.
+\end{remark}
+
+The axiom of choice has been the focus of particular investigation as it has been regarded as controversial for some of its consequences.
+While most mainstream mathematicians accept it, constructive mathematics rejects the axiom of choice as it does not lead to constructive proofs: it claims some objects exist even though they may not even be expressible or computable.
+It also implies the law of excluded middle (i.e.
+a proposition is either true or not), which constructive mathematics rejects.
+What does or does not require the axiom of choice is therefore ample subject of study.
+
+\begin{prop}
+	The following statements are equivalent:
+	\begin{itemize}
+		\item Axiom of choice: every set has a choice function
+		\item Well-ordering theorem: every set can be well ordered
+		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other
+		\item Zorn's lemma: every non-empty poset (i.e. partially ordered set) for which every chain (i.e. totally ordered subset) has an upper bound contains at least one maximal element
+		\item Every vector space has a basis
+	\end{itemize}
+\end{prop}
+
 \section{Numeric constructions}
-\begin{remark}
-	The constructions that follow exemplifies the standard way numbers are defines in mathematics and it is useful to gain certain insights.
-In the Assumptions of Physics, however, we prefer to understand numbers as order-theoretic concept (i.e.
-the integer numbers can be thought as a linearly ordered set where each element have an immediate successor and an immediate predecessor).
-\end{remark}
+In set theory, numbers are defined to be sets with particular properties.
+These constructions are instructive to gain insights in the foundations of mathematics.
+Note that, for the Assumptions of Physics, we found more productive to understand numbers as order-theoretic concept (i.e.
+the integer numbers can be thought as a linearly ordered set where each element has an immediate successor and an immediate predecessor).
+
+\subsection{Standard numbers}
+
 \begin{defn}\label{defn_successor_set}
@@
 The set of \textbf{natural numbers} $\mathbb{N}$ is the intersection of all successor sets.
-\end{defn}
-\begin{remark}
-	The natural numbers so constructed are known as von Neumann ordinals. They are as follows:
+The set of \textbf{positive naturals} is $\mathbb{N}^+ = \mathbb{N} \setminus \{ \emptyset \}$.
+\end{defn}
+\begin{remark}
+	The previous set exists only because one of the axioms tells us it does.
+The opposite axiom can be consistently chosen.
+The natural numbers so constructed are known as von Neumann ordinals.
+They are as follows:
 	\begin{itemize}
@@
 
-\section{Transfinite (i.e. infinite) numbers}
-\begin{remark}
+\begin{defn}
+	The \textbf{integers} can be defined from the natural number as an ordered pair of sign and value. That is, $\mathbb{Z} = \{ 0 \} \times \mathbb{N} \cup \{ 1 \} \times \mathbb{N}^+$.
+\end{defn}
+
+\begin{defn}
+	The \textbf{rationals} can be defined from the integers number as the equivalence class that leads to the same simplification.
+That is, let $Q = \mathbb{Z} \times \mathbb{Z}$ be all pairs of integers.
+Let $\sim \, \subset Q \times Q$ be an equivalence relation such that $(a_1, b_1) \sim (a_2, b_2)$ if and only if $a_1 b_2 = a_2 b_1$ (i.e.
+the pairs are proportional).
+The set of rationals is $\mathbb{Q} = Q_{/\sim}$.
+Given two integers $a, b \in \mathbb{Z}$, the rational number $\frac{a}{b}$ is the equivalence class $[(a,b)]_{\sim}$.
+\end{defn}
+
+\begin{defn}
+	The \textbf{reals} can be defined from the rationals as ``cuts'' that divide the rational in two contiguous sets.
+That is, let a real number be a partition $(A, B)$ such that $A \cup B = \mathbb{Q}$, $A \neq \emptyset$, $B \neq \emptyset$ and $a < b$ for all $a \in A$ and $b \in B$.
+The set of all reals $\mathbb{R}$ is the set of all such partitions.
+\end{defn}
+
+\begin{remark}
+	Operations such as additions, multiplication, ... can be defined appropriately as functions of the respective sets.
+\end{remark}
+
+\subsection{Ordinal numbers}
+
+To extend natural numbers beyond finite numbers we need to make a distinction.
 Cardinal numbers (i.e.
@@
 Cardinals and ordinals are in one to one correspondence only for finite values.
-\end{remark}
-
-\subsection{Ordinal numbers}
-
-\begin{remark}	
-	To define ordinals, we extend the notion of sequences to arbitrary levels of infinity using well-ordered sets.
-\end{remark}
-
-\begin{defn}[POS]
-	A \textbf{partially ordered set}, \marginleft{POS: $(A, \leq_A)$} or \textbf{POS}, is a set with an associated partial order.
+For infinite values, cardinals are a subset of the ordinals.
+
+\begin{remark}
+	To define ordinals, we extend the notion of sequences to arbitrary levels of infinity through well-ordered sets.
+\end{remark}
+
+\begin{defn}[Poset]
+	A \textbf{partially ordered set}, \marginleft{Poset: $(A, \leq_A)$} or \textbf{poset}, is a set with an associated partial order.
 That is, a tuple $(A, \leq_A)$ where $A$ is a set and $\leq_A$ is a partial order over $A$.
-A \textbf{linearly ordered set} is a POS whose order is linear.
+A \textbf{linearly ordered set} is a poset whose order is linear.
 \end{defn}
@@
 \begin{defn}[Initial segment]
-	The \textbf{initial segment} \marginleft{Initial segment} of a POS $(A, \leq_A)$ determined by $a \in A$ is the set of all elements before $a$. That is, $S_a=\{ x \in A \, | \, x < a \}$.
+	The \textbf{initial segment} \marginleft{Initial segment} of a poset $(A, \leq_A)$ determined by $a \in A$ is the set of all elements before $a$. That is, $S_a=\{ x \in A \, | \, x < a \}$.
 \end{defn}
@@
 \begin{defn}[Order morphisms]
-	Let $(A, \leq_A)$ \marginleft{Order homo/ \\ isomorphism} and $(B, \leq_B)$ be two POSs.
+	Let $(A, \leq_A)$ \marginleft{Order homo/ \\ isomorphism} and $(B, \leq_B)$ be two posets.
 An \textbf{order homomorphism} is a function $f : A \to B$ that preserves the ordering.
@@
 An \textbf{order isomorphism} is a bijective order homomorphism whose inverse is also an order homomorphism (i.e.
-the POSs have identical order).
+the posets have identical order).
 That is, $b_1 \leq_B b_2$ implies $f^{-1}(b_1) \leq_A f^{-1}(b_2)$.
@@
 	\end{itemize}
-	In a well-ordered set, every element always has an immediate successor, but not necessarily an infinite predecessor (e.g. $b_1$ in the third example).
+	In a well-ordered set, every element always has an immediate successor, but not necessarily an immediate predecessor (e.g. $b_1$ in the third example).
 \end{remark}
@@
 \begin{defn}[Ordinals]
-	An \textbf{ordinal} \marginleft{Ordinals: \ordinals} is a set $A$ that includes the previous and only the previous ordinals. That is, for all elements $a, b, c \in A$ (i.e. previous ordinals of $A$):
+	An \textbf{ordinal} \marginleft{Ordinals: \ordinals} is a set $A$ that includes the previous and only the previous ordinals (this makes $a \in A$ mean $a \ordless A$).
+That is, for all elements $a, b, c \in A$ (i.e.
+previous ordinals of $A$):
 	\begin{enumerate}
@@
 \begin{prop}[Ordinals are well-ordered]
-	All ordinals are well-ordered sets. That is, for each $A \in \ordinals$, $(A, \leq_A)$ is a well-ordered set where $a \leq b$ if either $a \in b$ or $a = b$.
+	All ordinals are well-ordered sets.
+That is, for each $A \in \ordinals$, $(A, \leq_A)$ is a well-ordered set where $a \leq b$ if either $a \in b$ or $a = b$.
+Moreover, the class $\ordinals$ is well ordered.
+That is, given a set of ordinals, one has lower ordinality than all others.
 \end{prop}
@@
 \begin{prop}[Well-orders are ordinals]
-	For every \marginleft{$\langle A \rangle$} well-ordered set $A$ there is a unique ordinal that fully characterizes its order.
-That is, there is a unique $\langle A \rangle \in \cardinals$ such that $a \ordeq A$.
+	For every \marginleft{$\langle A \rangle$, $\omega$} well-ordered set $A$ there is a unique ordinal that fully characterizes its order.
+That is, there is a unique $\langle A \rangle \in \cardinals$ such that $\langle A \rangle \ordeq A$.
 We say $\langle A \rangle$ is the \textbf{ordinality of $A$}.
+The ordinality of an infinite sequence is denote by $\omega$.
+That is, $\omega = (0, 1, 2, ...) = \langle \{x_i\}_{i=1}^\infty \rangle$ for all $\{x_i\}_{i=1}^\infty$.
 \end{prop}
@@
 	As an example, consider the von Neumann ordinal $3 = (0, 1, 2)$ as a well-ordered set.
-Similarly, the ordinality of the well-ordered set $(a_1, a_2, a_3)$ is $3 = (0, 1, 2)$ as share the same ordering.
+Similarly, the ordinality of the well-ordered set $(a_1, a_2, a_3)$ is $3 = (0, 1, 2)$ as they share the same ordering.
 The definitions are such that these ideas extend to infinite sets.
@@
 Let $(A, \leq_A)$ and $(B, \leq_B)$ be two well ordered sets.
-Let $(A \sqcup B, \leq)$ be the POS where:
+Let $(A \sqcup B, \leq)$ be the poset where:
 	\begin{itemize}
-		\item $(A, a_1) \leq (A, a_2)$ if $a_1 \leq_A a_2$
-		\item $(B, b_2) \leq (B, b_2)$ if $b_1 \leq_B b_2$
-		\item $(A, a) \leq (B, b)$ for all $a \in A$ and $b \in B$.  	
+		\item $(0, a_1) \leq (0, a_2)$ if $a_1 \leq_A a_2$
+		\item $(1, b_2) \leq (1, b_2)$ if $b_1 \leq_B b_2$
+		\item $(0, a) \leq (1, b)$ for all $a \in A$ and $b \in B$.  	
 	\end{itemize}
@@
 Let $(A, \leq_A)$  and $(B, \leq_B)$ be two well ordered sets.
-Let $(A \times B, \leq)$ be the POS where:
+Let $(A \times B, \leq)$ be the poset where:
 	\begin{itemize}
@@
 There is no upper limit to the ordinals.
-\end{remark}
+
+	Adding an ordinal of a lower power to the left does not change the ordinal.
+Adding finitely many elements at the beginning of an infinite sequence, still gives us an infinite sequence (i.e.
+$n + \omega = \omega$).
+In the same way, adding an infinite sequence before an infinite sequence of infinite sequences still gives us an infinite sequence of infinite sequences (i.e.
+$\omega +\omega^2 = \omega^2$).
+\end{remark}
+
+\begin{prop}[Cantor normal form]
+	Every ordinal number $\langle A \rangle$ can be written as a sum of powers of $\omega$.
+That is, $\langle A \rangle = \omega^{\alpha_1} c_1 + \omega^{\alpha_2} c_2 + ...
++ \omega^{\alpha_n} c_n$ where $n \in \mathbb{N}$, $c_i \in \mathbb{N}^+$ for all $1 \leq i \leq n$ and $\alpha_i \in \ordinals$ such that $\alpha_i > \alpha_j$ for all $i > j$.
+\end{prop}
 
@@
 That is, either $A \crdleq B$ or $B \crdleq A$.
-Additionally, there exists a bijective function between $A$ and $B$ only if there exists an injective function in both direction.
+Additionally, there exists a bijective function between $A$ and $B$ if and only if there exists an injective function in both direction.
 That is, $A \crdeq B$ if and only if $A \crdleq B$ and $B \crdleq A$.
@@
 That is, $a \ordleq b$ for all $b \in \ordinals$ such that $a \crdeq b$.
-The set of all cardinals is noted $\cardinals$ and, by definition $\cardinals \subset \ordinals$.
+The class of all cardinals is noted $\cardinals$ and, by definition $\cardinals \subset \ordinals$.
 \end{defn}
@@
 	For every \marginleft{$|A|$} set $A$ there is a unique cardinal that identifies the cardinality of $A$.
-That is, there exists a unique $|A| \in \cardinals$ such that $a \crdeq A$.
+That is, there exists a unique $|A| \in \cardinals$ such that $|A| \crdeq A$.
 We say $|A|$ is the \textbf{cardinality of $A$}.
@@
 \begin{prop}[Cantor's theorem]
-	A set has strictly lower cardinality than its power set. That is, $|A| < |\mathcal{P}(A)|$. If $a = |A|$ then we note $|\mathcal{P}(A)| = 2^a$
+	A set has strictly lower cardinality than its power set. That is, $|A| < |\mathcal{P}(A)|$. If $\alpha = |A|$ then we note $|\mathcal{P}(A)| = 2^\alpha$
 \end{prop}
@@
 \begin{prop}[Aleph numbers]
-	The cardinality \marginleft{Aleph \\numbers: $\aleph_a$} of infinite sets are well-ordered. They are noted $\aleph_a$ where $a \in \ordinals$.
+	The cardinality \marginleft{Aleph \\numbers: $\aleph_\alpha$} of infinite sets are well-ordered. They are noted $\aleph_\alpha$ where $\alpha \in \ordinals$.
 \end{prop}
@@
 \textbf{Continuum hypothesis}
-	There is \marginleft{Continuum hypothesis} not infinite set with cardinality in between that of the continuum and countable infinity. That is: $2^{\aleph_0} = \aleph_1$.
-
-\begin{remark}
-	The continuum hypothesis is proven to be independent on the axioms of set theory and is therefore an additional optional axiom.
+	There is \marginleft{Continuum hypothesis} no infinite set with cardinality in between that of the continuum and countable infinity. That is: $2^{\aleph_0} = \aleph_1$.
+
+\begin{remark}
+	The continuum hypothesis is proven to be independent of the axioms of set theory.
+It can be taken to be true or to be false without inconsistency.
 As the axiom deals with sets of cardinality lower than the continuum, this aspect of set theory may be interesting for those working on the Assumptions of Physics.
@@
 
-\section{Axiomatic set theory}
-
-Introduce paradoxes. Axiomatic set theory is a family of attempt to solve those paradoxes. There are different variations.
-
-Formalization to remove semantic paradoxes. All approaches of axiomatic set theory do that.
-
-%https://en.wikipedia.org/wiki/Zermelo%E2%80%93Fraenkel_set_theory
-
-%https://en.wikipedia.org/wiki/Von_Neumann–Bernays–Gödel_set_theory
-
-\subsection{Classes, sets and elements}
-
-As some collections cannot \marginright{TODO: diagram with Classes, proper classes and sets.} be put into other collections without creating paradoxes, the resolution of syntactic paradoxes is through the distinction between classes and sets.
-Classes are all those objects that can contain elements.
-Sets are particular classes that can also be elements of other classes.
-Therefore we can have the class of all sets, but we cannot have the set of all sets or the class of all classes.
-
-\subsection{Axiomatization}
-
-There are different approaches to the axiomatization of set theory with different degree of variation.
-One strategy, refered to as Zermelo–Fraenkel (ZFC), only defines sets.
-It avoids paradoxes by restricting set construction to subsets only.
-
-Another strategy, refered to as von Neumann–Bernays–Gödel (NBG), starts by defining classes and then defines sets as those classes that can be elements of classes.
-It avoids paradoxes by restricting set construction on predicates over elements.
-Since a proper class is not an element, if the definition creates a contradiction for an item $x$ (i.e.
-we have both $x \in S$ and $x \notin S$) than it must be that the element $x$ is a proper class.
-
-\textbf{Axioms of set theory}. Here are the axioms of a particular formulation of axiomatic set theory taken from CITE:
-\begin{enumerate}
-	\item Axiom of extent. If two classes have the same elements, then they are identical.
-	\item Axiom of class construction.
-Let $P(x)$ be a predicate expressed in terms of the symbols $\in, \OR, \AND, \NOT, \Rightarrow, \exists, \forall$, brackets, variables $x, y, z, A, B, ...$ Then there exists a class $C$ that consists of all the elements x which satisfy $P(x)$.
-	\item Every subclass of a set is a set.
-	\item $\emptyset$ is a set.
-	\item If $a$ and $b$ are sets, then $\{a, b\}$ is a set.
-	\item If $\mathcal{A}$ is a set of sets, then, $\bigcup\limits_{A \in \mathcal{A}} A$ is a set.
-	\item If $A$ is a set, the power set $\mathcal{P}(A)$ is a set.
-	\item Axiom of foundation. If $A \neq \emptyset$ is a set, there exists an element $a \in A$ such that $a \cap A = \emptyset$.
-	\item Axiom of replacement. If $A$ is a set and $f : A \to B$ is a surjective function, then $B$ is a set.
-	\item Axiom of choice. Every set has a choice function. That is, given a set $A$, there exists a function $f : \mathcal{P}(A) \setminus \{ \emptyset \} \to A$ such that $f(B) \in B$.
-	\item Axiom of infinity. There exists a successor set (as defined in \ref{defn_successor_set}).
-\end{enumerate}
-
-\begin{remark}
-	Axiom 1 serves to define class equality.
-Axiom 2 defines how to define classes based on predicates.
-Axioms 3-7 tells us that some classes, constructable by axiom 2, are sets.
-Axiom 8 guarantees that we cannot have an infinite sequences $\{ A_i \}_{i=1}^\infty$ such that $A_{i+1} \in A_i$.
-In particular, a set cannot be an element of itself.
-Axiom 9 tells us that any set that is ``smaller'' than a set (i.e.
-we can map another set onto it) is a set.
-Axiom 10 guarantees that we can always pick an element from a set, though it does not tell us how.
-Axiom 11 tells us that there are sets with infinite elements.
-Once the natural numbers are assumed to exist, higher cardinality of infinity can be constructed through, for example, the power set.
-
-	A lot of work in set theory has been and is to clarify the relationship between the axioms, in particular which are independent from the others.
-The axiom of choice has been controversial for some time and therefore has been one of the most studied.
-\end{remark}
-
-\begin{prop}
-	The following statements are equivalent:
-	\begin{itemize}
-		\item Axiom of choice: every set has a choice function
-		\item Well-ordering theorem: every set can be well ordered
-		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other
-		\item Zorn's lemma: every non-empty POS for which every chain (i.e. totally ordered subset) has an upper bound contains at least one maximal element
-		\item Every vector space has a basis
-	\end{itemize}
-\end{prop}
-
-\begin{remark}
-	It should be stressed that there are different approaches.
-Constructive set theory, as all constructive mathematics, rejects ``nonconstructive proofs'' in which the existence of an object is proved with no explicit algorithm as to how to constructive.
-This approach will, for example, exclude the axiom of choice.
-For those working on the Assumptions of Physics, then, it is important to understand that the foundations of math are not absolute.
-Therefore we have to understand the consequences of those choice and understand whether they are suitable for the foundations of physics.
-\end{remark}
-
 %\bibliographystyle{plain}

### 2021-05-19 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: adding references

@@
 \usepackage{tikz}
+\usetikzlibrary{positioning}
 \usepackage{hyperref}
@@
 
-TODO: cite https://www.math24.net/topics-set-theory and set theory book
+For more details on set theory, see for example \cite{settheory-tutorial, pinter2014book}.
 
@@
 
-A \textbf{class} \marginright{TODO: diagram with Classes, proper classes and sets.}  is any collection, therefore \textbf{sets are classes}.
+A \textbf{class} \marginright{ ~\\
+		\begin{tikzpicture}[scale = 0.3]
+		\node [draw, inner sep=6, rectangle] (class) {\small Class};
+		\node [below=.3 of class] (middle) {};
+		\node [below=1.3 of class] (lower) {};
+		\node [draw, inner sep=6, rectangle, align=center, left=0 of lower] (pclass) {\small Proper \\ class};
+		\node [draw, inner sep=6, rectangle, right=.2 of lower] (set) {\small Set};
+		\draw[->] (pclass.north) -| ++(0,.8) -|  (class.south);	
+		\draw[->] (set.north) -| ++(0,1.5) -|  (class.south);	
+		\draw[->] (class.east) -| ++(3,0) |-  (set.east);	
+		 %++(0,-1) -|
+		\node [left=-0.1 of middle] (isa) {\small is a};
+		\node at (4.1,.6) (contains) {\small contains};
+	\end{tikzpicture} 
+}  is any collection, therefore \textbf{sets are classes}.
 Some classes are not sets.
@@
 
-\textbf{Axioms of set theory}. Here are the axioms of a particular formulation of axiomatic set theory taken from CITE that follows NBG:
+\textbf{Axioms of set theory}. Here \marginleft{Set theory axioms} are the axioms of a particular formulation of axiomatic set theory taken from \cite{pinter2014book} that follows NBG:
 \begin{enumerate}
@@
 
-%\bibliographystyle{plain}
-%\bibliography{bibliography}
+\bibliographystyle{plain}
+\bibliography{bibliography}
 

### 2021-05-26 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: Christine's review

@@
 
-Set theory is important for those working on the Assumptions of Physics for at least three reasons.
+Set theory is important for those working on Assumptions of Physics for at least three reasons.
 First, it is a foundational framework in mathematics.
@@
 
-There are two rough branches of set theory: naive and axiomatic.
+There are roughly two branches of set theory: naive and axiomatic.
 Naive set theory is built on intuitive concepts, and as such is not fully formalized and is open to potential paradoxes.
@@
 
-For more details on set theory, see for example \cite{settheory-tutorial, pinter2014book}.
+For more details on set theory, see for example \cite{pinter2014book, settheory-tutorial}.
 
@@
 We take sets and elements as primitive objects, with no formal definition.
+This is the only difference between naive and axiomatic set theory, therefore all the formal definitions that follow apply to both.
 A \textbf{set} $A$ is a collection of elements.
@@
 predicate) $P(e)$, and is noted $A = \{e \, | \,  P(e) \}$.
-The rule may include symbols like for all $\forall$, exists $\exists$, logical connectors AND $\AND$, OR $\OR$ and $\NOT$.
+The rule may include symbols like for all $\forall$, exists $\exists$, logical connectors AND $\AND$, OR $\OR$ and NOT $\NOT$.
 
@@
 	 \marginleft{Common sets: $\emptyset$, $\{a\}$, $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$, $\mathbb{C}$}
-	  $\emptyset$ is the set with no elements. A \textbf{singleton} is the set of a single value $\{a\}$.
+	  $\emptyset$ is the set with no elements. A \textbf{singleton} is the set of a single element $\{a\}$.
 	 The \textbf{set of natural numbers} $\mathbb{N} = \{ 0, 1, 2, ...
@@
 	Let $\mathcal{A} \subseteq \mathcal{P}(A)$ be a set of subsets of $A$.
-The operations of union $\bigcup\limits_{A \in \mathcal{A}} A$ and intersection $\bigcap\limits_{A \in \mathcal{A}} A$ can be extend to span all elements.
+The operations of union $\bigcup\limits_{A \in \mathcal{A}} A$ and intersection $\bigcap\limits_{A \in \mathcal{A}} A$ can be extended to span all elements.
 \end{defn}
@@
 That is $(a, b) \neq (b, a)$.
-This can be constructed setting $(a, b) = \{ \{a\}, \{a, b\}\}$.
+This can be constructed by setting $(a, b) = \{ \{a\}, \{a, b\}\}$.
 \end{defn}
@@
 The disjoint union $A \sqcup B$ is the union where the elements of the sets are always treated as distinct.
-That is, given $A \sqcup B = \{0\} \times A \cup \{1\} \times B$.
-\end{defn}
-
-\begin{remark}
-	Ordered pairs and disjoint unions exemplify the effort in set theory to construct \emph{all object} starting from the primitives of set and set inclusion.
+That is, given $A \sqcup B = (\{0\} \times A ) \cup ( \{1\} \times B )$.
+\end{defn}
+
+\begin{remark}
+	Ordered pairs and disjoint unions exemplify the effort in set theory to construct \emph{all objects} starting from just two primitives: set and set inclusion.
+The downside is that the constructions do not map exactly to our intuition.
+For example, we still think an ordered pair as a set, so we want to write $b \in (a, b)$, though, looking at the definition $b$ itself is not directly a member of $(a, b)$.
 \end{remark}
@@
 	Relations provide a common foundation for functions (i.e.
-the relation is the set of pairs $(x, f(x))$), equivalences and order (i.e.
+the relation is the set of pairs $(x, f(x))$ ), equivalences and order (i.e.
 the relation is the set of pairs $(a, b)$ such that $a=b$ or $a\leq b$ respectively).
@@
 If $A=B$ (i.e.
-domain and codomain coincide) the relation is said \textbf{homogeneous}, \textbf{heterogeneous} if not.
+domain and codomain coincide) the relation is said to be \textbf{homogeneous}, \textbf{heterogeneous} if not.
 \end{defn}
@@
 	If $X \subseteq A$, \marginleft{Image and\\ preimage} the \textbf{image of $X$ under $R$} is the set of all elements in $B$ that are related to at least one element in $X$.
-The \textbf{image of $R$} is set of all elements in $B$ that are related to at least one element in $A$ (i.e.
+The \textbf{image of $R$} is the set of all elements in $B$ that are related to at least one element in $A$ (i.e.
 the image of the full set $A$).
@@
 	If $Y \subseteq B$, the \textbf{preimage of $Y$ under $R$} is the set of all elements in $A$ that are related to at least one element in $Y$.
-The \textbf{preimage of $R$} is set of all elements in $A$ that are related to at least one element in $B$ (i.e.
+The \textbf{preimage of $R$} is the set of all elements in $A$ that are related to at least one element in $B$ (i.e.
 the preimage of the full set $B$).
@@
 In this case, the preimage of $R$ is the whole $A$.
-		\item[Surjective or right-total.] Every element $b \in B$ is related to at least one element $a \in B$.
+		\item[Surjective or right-total.] Every element $b \in B$ is related to at least one element $a \in A$.
 That is, for each $b \in B$ there exists at least one $a \in A$ such that $aRb$.
@@
 \begin{prop}[Properties of image/preimage functions]
-	The induced relations $f : \mathcal{P}(A) \to \mathcal{P}(B)$ and $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ are functions and satisfy the following properties:
+	The induced relations $f : \mathcal{P}(A) \to \mathcal{P}(B)$ and $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ are functions and satisfy the following properties for all $\mathcal{A} \subseteq \mathcal{P}(A)$ and $\mathcal{B} \subseteq \mathcal{P}(B)$:
 	\begin{enumerate}
-		\item $f(\bigcup\limits_{A \in \mathcal{A}} A) = \bigcup\limits_{A \in \mathcal{A}} f(A)$
-		\item $f^{-1}(\bigcup\limits_{A \in \mathcal{A}} A) = \bigcup\limits_{A \in \mathcal{A}} f^{-1}(A)$
-		\item $f^{-1}(\bigcap\limits_{A \in \mathcal{A}} A) = \bigcap\limits_{A \in \mathcal{A}} f^{-1}(A)$
+		\item $f(\bigcup\limits_{U \in \mathcal{A}} U) = \bigcup\limits_{U \in \mathcal{A}} f(U)$
+		\item $f^{-1}(\bigcup\limits_{V \in \mathcal{B}} V) = \bigcup\limits_{V \in \mathcal{B}} f^{-1}(V)$
+		\item $f^{-1}(\bigcap\limits_{V \in \mathcal{B}} V) = \bigcap\limits_{V \in \mathcal{B}} f^{-1}(V)$
 	\end{enumerate}
@@
 Also note that the image of the intersection is not in general the intersection of the image, and how maps that preserve set theoretic structures (e.g.
-topologies and $\sigma$-algebras) are defined based on reverse images.
+topologies and $\sigma$-algebras) are defined based on preimages.
 \end{remark}
@@
 \begin{defn}[Fiber]
-	Given \marginleft{Fiber} a function $f : A \to B$, the A \textbf{fiber of $b \in B$ under $f$} is the subset of $A$ of where the function takes the value $b$.
-That is, the reverse image $f^{-1}(\{b\})$ of the singleton.
+	Given \marginleft{Fiber} a function $f : A \to B$, the \textbf{fiber of $b \in B$ under $f$} is the subset of $A$ of where the function takes the value $b$.
+That is, the preimage $f^{-1}(\{b\})$ of the singleton.
 Fibers over real functions are also called \textbf{level sets}.
@@
 \begin{defn}[Restriction]
-	Let $f : A  \to B$ \marginleft{Restriction: $f|_C$} be a function and let $C \subseteq B$. Then the \textbf{restriction of $f$ over $C$} is the function $f|_C : C \to B$ such that $f|_C(c) = f(c)$.
+	Let $f : A  \to B$ \marginleft{Restriction: $f|_C$} be a function and let $C \subseteq A$.
+Then the \textbf{restriction of $f$ over $C$} is the function $f|_C : C \to B$ such that $f|_C(c) = f(c)$ for all $c \in C$.
 \end{defn}
@@
 		\item If $f : A \to B$ and $g : B \to C$ are (partial) functions, then $g \circ f$ is a (partial) function.
-		\item Let $f : A \to B$ is a bijective function, then $f \circ f^{-1} = \Id_B$ and $f^{-1} \circ f = \Id_A$.
+		\item Let $f : A \to B$ be a bijective function, then $f \circ f^{-1} = \Id_B$ and $f^{-1} \circ f = \Id_A$.
 	\end{itemize}
@@
 	The notation for function sets is due to the cardinality.
-If $A$ has $n$ elements and $|B|$ has $m$ elements, each function is given by choosing an element of $B$ for each element of $A$, that is $m$ choices $n$ times, that is $m^n$.
-The power set is noted $2^A$ because each subset can be identified by a function $U : A \to \{0 , 1\}$, where $f(a)=0$ corresponds to $a \notin U$ and $f(a)=0$ corresponds to $b \in U$.
+If $A$ has $n$ elements and $B$ has $m$ elements, each function is given by choosing an element of $B$ for each element of $A$, that is $m$ choices $n$ times, that is $m^n$.
+The power set is noted $2^A$ because each subset $U \subseteq A$ can be identified by a unique function $f : A \to \{0 , 1\}$, where for all $a\in A$ then $f(a)=0$ corresponds to $a \notin U$ and $f(a)=1$ corresponds to $a \in U$.
 \end{remark}
@@
 \times A_n \to B$ be an $n$-ary function.
-We say a set \textbf{$A$ is closed under $f$} if whenever the function is applied to elements of $A$ returns an element of $A$.
-That is, $A \subseteq A_i$ for all $1\leq i \leq n$, $A \subset B$ and $f(A, A, ..., A) \subseteq A$.
+We say a set \textbf{$A$ is closed under $f$} if whenever the function is applied to elements of $A$ it returns an element of $A$.
+That is, $A \subseteq A_i$ for all $1\leq i \leq n$, $A \subseteq B$ and $f(A, A, ..., A) \subseteq A$.
 \end{defn}
@@
 	For example, integers are closed under addition, subtraction and multiplication but are not closed under division.
-Most mathematical structures (groups, topologies, $\sigma$-algebra, vector spaces, ...) define sets closed under some operations.
+Most mathematical structures (groups, topologies, $\sigma$-algebras, vector spaces, ...) define sets closed under some operations.
 \end{remark}
@@
 	An \textbf{indexed family}, \marginleft{Families, sequences: $\{x_i\}_{i \in I}$, $\{x_i\}_{i=1}^{\infty}$} or simply a \textbf{family}, $\{x_i\}_{i \in I}$ is a collection of elements identified by an \textbf{index set} $I$.
-More formally, an indexed family is a tuple $\langle X, I, x \rangle$ where $X$ and $I$ are sets and $x: I \to X$ is a function such that $i \mapsto x_i$.
-A \textbf{sequence} is a family for which the index set is a continuous interval of natural numbers.
+More formally, an indexed family is a tuple $( X, I, x )$ where $X$ and $I$ are sets and $x: I \to X$ is a function such that $i \mapsto x_i$.
+A \textbf{sequence} is a family for which the index set is a contiguous interval of natural numbers.
 A \textbf{finite sequence} of $n$ elements is noted $\{x_i\}_{i=1}^{n}$ while an \textbf{infinite sequence} is noted $\{x_i\}_{i=1}^{\infty}$.
@@
 \begin{defn}[Set-theoretic limit]
-	Let $\{A_i\}_{i=0}^{\infty}$ \marginleft{Set limit: \\ $\lim\limits_{i \to \infty} A_i = A$} be a sequence of sets.
+	Let $\{A_i\}_{i=1}^{\infty}$ \marginleft{Set limit: \\ $\lim\limits_{i \to \infty} A_i = A$} be a sequence of sets.
 We define the \textbf{limit infimum} as $\liminf\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1}\bigcap\limits_{j \geq i} A_j$ and the \textbf{limit supremum} as $\limsup\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1}\bigcup\limits_{j \geq i} A_j$.
@@
 \begin{defn}[Monotone sequence]
-	A \textbf{monotone sequence} \marginleft{Monotone sequence: \\ $A_1 \subseteq A_2 \subseteq ...$\\ $A_1 \supseteq A_2 \supseteq ...$} is a sequence of sets $\{A_i\}_{i=0}^{\infty}$ where each set is a subset (or superset) of the preceding.
+	A \textbf{monotone sequence} \marginleft{Monotone sequence: \\ $A_1 \subseteq A_2 \subseteq ...$\\ $A_1 \supseteq A_2 \supseteq ...$} is a sequence of sets $\{A_i\}_{i=1}^{\infty}$ where each set is a subset (or superset) of the preceding.
 More specifically, we distinguish four cases:
 	\begin{description}
-		\item[Increasing] when $A_i \subseteq A_{i+1}$ for all $i>0$
-		\item[Decreasing] when $A_i \supseteq A_{i+1}$ for all $i>0$
-		\item[Strictly increasing] when $A_i \subset A_{i+1}$ for all $i>0$
-		\item[Strictly decreasing] when $A_i \supset A_{i+1}$ for all $i>0$
+		\item[Increasing] when $A_i \subseteq A_{i+1}$ for all $i \geq 1$
+		\item[Decreasing] when $A_i \supseteq A_{i+1}$ for all $i \geq 1$
+		\item[Strictly increasing] when $A_i \subset A_{i+1}$ for all $i \geq 1$
+		\item[Strictly decreasing] when $A_i \supset A_{i+1}$ for all $i \geq 1$
 	\end{description}
@@
 \begin{remark}
-	We will cover the notions of orders that are necessary for the definition of ordinals and cardinal numbers. Orders will be treated more in detail in a Bare Minimum dedicated to order theory.
+	We will cover the notions of orders that are necessary for the definition of ordinal and cardinal numbers. Orders will be treated more in detail in a Bare Minimum dedicated to order theory.
 \end{remark}
@@
 		\item[Symmetric.] If $a$ is related to $b$, then $b$ is related to $a$. That is, if $aRb$ then $bRa$.
-		\item[Antisymmetric.] If $a$ is related to $b$ and $b$ is related to $a$, then $a$ and $b$ are the same element. That is, if $aRb$ and $aRb$ then $a=b$.
+		\item[Antisymmetric.] If $a$ is related to $b$ and $b$ is related to $a$, then $a$ and $b$ are the same element. That is, if $aRb$ and $bRa$ then $a=b$.
 		\item[Transitive.] If $a$ is related to $b$ and $b$ is related to $c$, then $a$ is related to $c$. That is, if $aRb$ and $bRc$ then $aRc$.
@@
 \begin{defn}[Orders and equivalence]
-	A \textbf{partial order} (noted $\leq$, $\lesssim$, $\preceq$) \marginleft{Order, equivalence: $\leq$, $<$, $\equiv$} is a homogeneous binary relationship that is reflexive, antisymmetric and transitive.
+	A \textbf{partial order} \marginleft{Order, equivalence: $\leq$, $<$, $\equiv$} (noted for example $\leq$, $\lesssim$, $\preceq$)  is a homogeneous binary relationship that is reflexive, antisymmetric and transitive.
 A \textbf{linear order} or \textbf{total order} is an order that is total.
-An order is \textbf{strict} (noted $<$, $\prec$) if it is irreflexive instead of reflexive.
-
-	An \textbf{equivalence relation} (noted $\equiv$, $\sim$) is a homogeneous binary relationship that is reflexive, symmetric and transitive.
+An order is \textbf{strict} (noted for example $<$, $\prec$) if it is irreflexive instead of reflexive.
+
+	An \textbf{equivalence relation} (noted for example $\equiv$, $\sim$) is a homogeneous binary relationship that is reflexive, symmetric and transitive.
 \end{defn}
@@
 \begin{defn}[Partitions]
-	A \textbf{partition} \marginleft{Partitions} of a set $A$ is a collections $P \subset \mathcal{P}(A)$ of disjoint subsets that cover all of $A$. That is:
+	A \textbf{partition} \marginleft{Partitions} of a set $A$ is a collection $P \subset \mathcal{P}(A)$ of disjoint subsets that cover all of $A$. That is:
 	\begin{enumerate}
@@
 The \textbf{equivalence class of $a$ by $\sim$} is the set $[a]_{\sim} = a_{/\sim} = \{b \in A \, | \, b \sim a \}$ of all the elements of $A$ equivalent to $a$.
-The \textbf{quotient set of $A$ by $\sim$} $A_{/\sim} = \{ a_{/\sim} \, | \, a \in A \}$  is the set of all of the equivalences classes by $\sim$.
-\end{defn}
+The \textbf{quotient set of $A$ by $\sim$} $A_{/\sim} = \{ a_{/\sim} \, | \, a \in A \}$  is the set of all of the equivalence classes by $\sim$.
+\end{defn}
+
+TODO: diagram
 
@@
 
-This is an important insight for those working on the Assumptions of Physics: the foundations of mathematics are not absolute.
-We have to be conscious of the consequences of those choice and understand whether they are suitable for the foundations of physics.
+This is an important insight for those working on Assumptions of Physics: the foundations of mathematics are not absolute.
+We have to be conscious of the consequences of those choices and understand whether they are suitable for the foundations of physics.
 
@@
 While they are generally called paradoxes, the following are really contradictions and that is why axiomatic set theory needs to avoid them.
-The first type of paradoxes to naive set theory arise from the use of natural language and the meaning of the words, typically when this is self-referential.
+The first type of paradoxes for naive set theory arises from the use of natural language and the meaning of the words, typically when this is self-referential.
 
@@
 
-These type of paradoxes are solved by restricting to symbolic expressions of a formal language, such as predicate logic.
+This type of paradoxes is solved by restricting to symbolic expressions of a formal language, such as predicate logic.
 The consequence is that, at the foundational level, mathematics does not keep track of what the symbols mean.
 
-The first type of paradoxes are purely logical, so they can be constructed in the formal system.
+The second type of paradoxes is purely logical, so they can be constructed in the formal system.
 
 \textbf{Russell's paradox}.
-Let $R$ be the set of all sets that contain themselves.
+Let $R$ be the set of all sets that do not contain themselves.
 That is, $R=\{ x \, | \, x \notin x\}$.
@@
 Note that axiomatic set theory does not distinguish between objects that can or cannot contain other objects.
-All elements are classes, all element can, in principle, contain something.
+All elements are classes, all elements can, in principle, contain something.
 In fact, since equality of two objects coincides on whether they have the same elements, two objects must contain something different to be different.
@@
 As a consequence, the proposition $3 \in 4$ is valid, and happens to be true if $3$ and $4$ are naturals, though it is not true if $3$ and $4$ are reals.
-From the perspective of the Assumptions of Physics, this is not great as the framework allows propositions that are not physically meaningful.
-Radical different approaches to the foundations of mathematics (e.g.
+From the perspective of Assumptions of Physics, this is not great as the framework allows propositions that are not physically meaningful.
+Radically different approaches to the foundations of mathematics (e.g.
 through category theory) do not exhibit this particular problem, though they exhibit others (e.g.
@@
 One strategy for the axiomatization of set theory, refered to as Zermelo–Fraenkel (ZFC), only defines sets.
+The C in ZFC stands for the addition of the axiom of choice.
 It avoids paradoxes by restricting set construction to subsets only.
@@
 Since a proper class cannot be an element, if the definition creates a contradiction for an item $x$ (i.e.
-we have both $x \in S$ and $x \notin S$) than it must be that the element $x$ is a proper class.
+we have both $x \in S$ and $x \notin S$) then it must be that the element $x$ is a proper class.
 
@@
 Let $P(x)$ be a predicate expressed in terms of the symbols $\in, \OR, \AND, \NOT, \Rightarrow, \exists, \forall$, brackets, variables $x, y, z, A, B, ...$ Then there exists a class $C$ that consists of all the elements (i.e.
-sets) x which satisfy $P(x)$.
+sets) $x$ which satisfy $P(x)$.
 	\item Every subclass of a set is a set.
@@
 	\item Axiom of replacement. If $A$ is a set and $f : A \to B$ is a surjective function, then $B$ is a set.
-	\item Axiom of choice. Every set has a choice function. That is, given a set $A$, there exists a function $f : \mathcal{P}(A) \setminus \{ \emptyset \} \to A$ such that $f(B) \in B$.
-	\item Axiom of infinity. There exists a successor set (as defined in \ref{defn_successor_set}).
+	\item Axiom of choice.
+Every set has a choice function.
+That is, given a set $A$, there exists a function $f : \mathcal{P}(A) \setminus \{ \emptyset \} \to A$ such that for all non-empty $B \subseteq A$ we have $f(B) \in B$.
+	\item Axiom of infinity. There exists a successor set (as defined in Definition \ref{defn_successor_set}).
 \end{enumerate}
@@
 Axiom 2 defines how to define classes based on predicates.
-Axioms 3-7 tells us that some classes, constructable by axiom 2, are sets.
-Axiom 8 guarantees that we cannot have an infinite sequences $\{ A_i \}_{i=1}^\infty$ such that $A_{i+1} \in A_i$.
+Axioms 3-7 tell us that some classes, constructable by axiom 2, are sets.
+Axiom 8 guarantees that we cannot have infinite sequences $\{ A_i \}_{i=1}^\infty$ such that $A_{i+1} \in A_i$.
 In particular, a set cannot be an element of itself.
@@
 we can map another set onto it) is a set.
-Axiom 10 guarantees that we can always pick an element from a set, though it does not tell us how.
+Axiom 10, the axiom of choice, guarantees that we can always pick an element from a set, though it does not tell us how.
 Axiom 11 tells us that there are sets with infinite elements.
@@
 		\item Axiom of choice: every set has a choice function
-		\item Well-ordering theorem: every set can be well ordered
-		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other
+		\item Well-ordering theorem: every set can be well ordered (defined below)
+		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other (i.e. $|A| < |B|$, $|A| = |B|$ or $|A| > |B|$)
 		\item Zorn's lemma: every non-empty poset (i.e. partially ordered set) for which every chain (i.e. totally ordered subset) has an upper bound contains at least one maximal element
@@
 In set theory, numbers are defined to be sets with particular properties.
-These constructions are instructive to gain insights in the foundations of mathematics.
-Note that, for the Assumptions of Physics, we found more productive to understand numbers as order-theoretic concept (i.e.
-the integer numbers can be thought as a linearly ordered set where each element has an immediate successor and an immediate predecessor).
+These constructions are instructive to gain insights into the foundations of mathematics.
+Note that, for Assumptions of Physics, we found it more productive to understand numbers as an order-theoretic concept (i.e.
+the integer numbers can be thought of as a linearly ordered set where each element has an immediate successor and an immediate predecessor).
 
@@
 \begin{remark}
-	The previous set exists only because one of the axioms tells us it does.
-The opposite axiom can be consistently chosen.
+	The set of the natural numbers exists only because the axiom of infinity tells us it does.
+The negation of the axiom of infinity can be consistently chosen, which would lead to a different version of set theory.
 The natural numbers so constructed are known as von Neumann ordinals.
@@
 \begin{defn}
-	The \textbf{integers} can be defined from the natural number as an ordered pair of sign and value. That is, $\mathbb{Z} = \{ 0 \} \times \mathbb{N} \cup \{ 1 \} \times \mathbb{N}^+$.
+	The \textbf{integers} can be defined from the natural numbers as an ordered pair of sign and value. That is, $\mathbb{Z} = (\{ 0 \} \times \mathbb{N}) \cup (\{ 1 \} \times \mathbb{N}^+)$.
 \end{defn}
@@
 \begin{defn}
-	The \textbf{rationals} can be defined from the integers number as the equivalence class that leads to the same simplification.
+	The \textbf{rationals} can be defined from the integer numbers as the equivalence class that leads to the same simplification.
 That is, let $Q = \mathbb{Z} \times \mathbb{Z}$ be all pairs of integers.
@@
 \begin{defn}
-	The \textbf{reals} can be defined from the rationals as ``cuts'' that divide the rational in two contiguous sets.
+	The \textbf{reals} can be defined from the rationals as ``cuts'' that divide the rationals in two contiguous sets.
 That is, let a real number be a partition $(A, B)$ such that $A \cup B = \mathbb{Q}$, $A \neq \emptyset$, $B \neq \emptyset$ and $a < b$ for all $a \in A$ and $b \in B$.
@@
 \begin{remark}
-	Operations such as additions, multiplication, ... can be defined appropriately as functions of the respective sets.
+	Operations such as additions, multiplications, ... can be defined appropriately as functions of the respective sets.
 \end{remark}
@@
 	To define ordinals, we extend the notion of sequences to arbitrary levels of infinity through well-ordered sets.
+Note that we introduced the non-standard symbols $\ordleq$, $\ordeq$ and $\langle A \rangle$ to make the notation more compact.
 \end{remark}
@@
 That is, for every $B \subseteq A$, there exists an $a \in B$ such that $a \leq_A b$ for all $b \in B$.
-Let $A$ and $B$ be two well ordered set, we say $A$ \textbf{has lower ordinality than} $B$, noted $A \ordleq B$, if $A$ is orderwise identical to the beginning of $B$.
+Let $A$ and $B$ be two well-ordered sets.
+We say $A$ \textbf{has lower or equal ordinality than} $B$, noted $A \ordleq B$, if $A$ is orderwise identical to the beginning of $B$.
 That is, $A$ is order isomorphic to an initial segment of $B$.
@@
 \begin{prop}[Ordinality]
-	If $A$ and $B$ are two well ordered sets, one of them has lower cardinality than the other.
+	If $A$ and $B$ are two well-ordered sets, one of them has lower or equal ordinality than the other.
 That is, either $A \ordleq B$ or $B \ordleq A$.
@@
 	All ordinals are well-ordered sets.
-That is, for each $A \in \ordinals$, $(A, \leq_A)$ is a well-ordered set where $a \leq b$ if either $a \in b$ or $a = b$.
+That is, for each $A \in \ordinals$, $(A, \leq_A)$ is a well-ordered set where $a \leq_A b$ if either $a \in b$ or $a = b$.
 Moreover, the class $\ordinals$ is well ordered.
-That is, given a set of ordinals, one has lower ordinality than all others.
+That is, given a set of ordinals, one has lower or equal ordinality than all others.
 \end{prop}
@@
 	For every \marginleft{$\langle A \rangle$, $\omega$} well-ordered set $A$ there is a unique ordinal that fully characterizes its order.
-That is, there is a unique $\langle A \rangle \in \cardinals$ such that $\langle A \rangle \ordeq A$.
+That is, there is a unique $\langle A \rangle \in \ordinals$ such that $\langle A \rangle \ordeq A$.
 We say $\langle A \rangle$ is the \textbf{ordinality of $A$}.
-The ordinality of an infinite sequence is denote by $\omega$.
+The ordinality of an infinite sequence is denoted by $\omega$.
 That is, $\omega = (0, 1, 2, ...) = \langle \{x_i\}_{i=1}^\infty \rangle$ for all $\{x_i\}_{i=1}^\infty$.
@@
 	The addition \marginleft{Ordinal addition: $\langle A \rangle + \langle B \rangle$} of two ordinals is defined by concatenating the first order with the second.
-Let $(A, \leq_A)$ and $(B, \leq_B)$ be two well ordered sets.
+Let $(A, \leq_A)$ and $(B, \leq_B)$ be two well-ordered sets.
 Let $(A \sqcup B, \leq)$ be the poset where:
@@
 		\item $(0, a_1) \leq (0, a_2)$ if $a_1 \leq_A a_2$
-		\item $(1, b_2) \leq (1, b_2)$ if $b_1 \leq_B b_2$
+		\item $(1, b_1) \leq (1, b_2)$ if $b_1 \leq_B b_2$
 		\item $(0, a) \leq (1, b)$ for all $a \in A$ and $b \in B$.  	
@@
 \begin{remark}
-	Addition in non-commutative (i.e.
+	Addition is non-commutative (i.e.
 $\langle A \rangle + \langle B \rangle \neq \langle B \rangle + \langle A \rangle$) for infinite ordinals.
@@
 	The multiplication \marginleft{Ordinal multiplication: $\langle A \rangle \langle B \rangle$} of two ordinals is defined by ordering their Cartesian product.
-Let $(A, \leq_A)$  and $(B, \leq_B)$ be two well ordered sets.
+Let $(A, \leq_A)$  and $(B, \leq_B)$ be two well-ordered sets.
 Let $(A \times B, \leq)$ be the poset where:
@@
 	Adding an ordinal of a lower power to the left does not change the ordinal.
-Adding finitely many elements at the beginning of an infinite sequence, still gives us an infinite sequence (i.e.
+Adding finitely many elements at the beginning of an infinite sequence still gives us an infinite sequence (i.e.
 $n + \omega = \omega$).
@@
 how many elements it contains) is defined in terms of injection/bijection: a set is smaller than another if it can be injected into it, and it is as big as another if it can be put into a one-to-one correspondence.
+Note that we introduced the non-standard symbols $\crdleq$ and $\crdeq$ to make the notation more compact.
 \end{remark}
@@
 	Let $A$ \marginleft{Cardinality} and $B$ be two sets.
-If there exists a injective function $f : A \to B$ then we say $A$ \textbf{has lower cardinality than} $B$, noted $A \crdleq B$.
+If there exists a injective function $f : A \to B$ then we say $A$ \textbf{has lower or equal cardinality than} $B$, noted $A \crdleq B$.
 If there exists a bijective function we say they have \textbf{the same cardinality}, noted $A \crdeq B$.
@@
 \begin{prop}[Cardinal comparability of sets]
-	Given two sets $A$ and $B$, one must have lower cardinality than the other.
+	Given two sets $A$ and $B$, one must have lower or equal cardinality than the other.
 That is, either $A \crdleq B$ or $B \crdleq A$.
-Additionally, there exists a bijective function between $A$ and $B$ if and only if there exists an injective function in both direction.
+Additionally, there exists a bijective function between $A$ and $B$ if and only if there exists an injective function in both directions.
 That is, $A \crdeq B$ if and only if $A \crdleq B$ and $B \crdleq A$.
@@
 \begin{defn}[Cardinal number]
-	A \textbf{cardinal}  \marginleft{Cardinals: $\cardinals$} is an ordinal $a \in \ordinals$ that has lower ordinality of any ordinal that has the same cardinality.
+	A \textbf{cardinal}  \marginleft{Cardinals: $\cardinals$} is an ordinal $a \in \ordinals$ that has lower or equal ordinality than any ordinal that has the same cardinality.
 That is, $a \ordleq b$ for all $b \in \ordinals$ such that $a \crdeq b$.
@@
 \begin{prop}[Aleph numbers]
-	The cardinality \marginleft{Aleph \\numbers: $\aleph_\alpha$} of infinite sets are well-ordered. They are noted $\aleph_\alpha$ where $\alpha \in \ordinals$.
+	The infinite cardinals  \marginleft{Aleph \\numbers: $\aleph_\alpha$} are well-ordered. They are noted $\aleph_\alpha$ where $\alpha \in \ordinals$.
 \end{prop}
@@
 		\item real $\mathbb{R}$ and complex $\mathbb{C}$ numbers
-		\item real and complex euclidean spaces: $\mathbb{R}^n$, $\mathbb{C}^n$
-		\item continuous function over euclidean spaces: $C(\mathbb{R}^n, \mathbb{R}^m)$
+		\item real and complex Euclidean spaces: $\mathbb{R}^n$, $\mathbb{C}^n$
+		\item continuous functions over Euclidean spaces: $C(\mathbb{R}^n, \mathbb{R}^m)$
 		\item standard topology and Borel algebra on $\mathbb{R}^n$ (i.e. the set of open sets and Borel set respectively)
@@
 This is not a coincidence.
-In science we are interested in object that can be experimentally identified.
+In science we are interested in objects that can be experimentally identified.
 Every experimental test can acquire finite information in finite time and, assuming arbitrarily long time, we can run at most countably many tests.
@@
 Even if nature allows for wider variety, it would be merged into equivalence classes of objects that are experimentally indistinguishable.
-Much of issues of higher cardinals can therefore be safely set aside by those working on the Assumptions of Physics.
+Many of the issues of higher cardinals can therefore be safely set aside by those working on Assumptions of Physics.
 \end{remark}
@@
 \begin{remark}
-	 While it can be proven that the cardinality of the continuum $2^{\aleph_0}$, yet which one is undetermined.
+	 While it can be proven that the cardinality of the continuum $2^{\aleph_0}$ is an aleph number, which one is undetermined in standard set theory.
 \end{remark}
@@
 \textbf{Continuum hypothesis}
-	There is \marginleft{Continuum hypothesis} no infinite set with cardinality in between that of the continuum and countable infinity. That is: $2^{\aleph_0} = \aleph_1$.
+	There is \marginleft{Continuum hypothesis} no infinite set with cardinality in between that of the continuum and countable infinity. That is, $2^{\aleph_0} = \aleph_1$.
 
@@
 It can be taken to be true or to be false without inconsistency.
-As the axiom deals with sets of cardinality lower than the continuum, this aspect of set theory may be interesting for those working on the Assumptions of Physics.
+As the axiom deals with sets of cardinality lower than the continuum, this aspect of set theory may be interesting for those working on Assumptions of Physics.
 It could be, in principle, that the distinguishable configurations in space-time have a cardinality that is neither countable nor that of the continuum.

### 2021-05-26 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: added image for partitions

@@
 The downside is that the constructions do not map exactly to our intuition.
-For example, we still think an ordered pair as a set, so we want to write $b \in (a, b)$, though, looking at the definition $b$ itself is not directly a member of $(a, b)$.
+For example, we still think as an ordered pair as a set, so we want to write $b \in (a, b)$, though, looking at the definition $b$ itself is not directly a member of $(a, b)$.
 \end{remark}
@@
 \subsection{Relations and functions}
-
-\begin{remark}
-	Relations provide a common foundation for functions (i.e.
-the relation is the set of pairs $(x, f(x))$ ), equivalences and order (i.e.
-the relation is the set of pairs $(a, b)$ such that $a=b$ or $a\leq b$ respectively).
-This exemplifies the effort in set theory to find general structures.
-\end{remark}
 
@@
 }
+\begin{remark}
+	Relations provide a common foundation for functions (i.e.
+the relation is the set of pairs $(x, f(x))$ ), equivalences and order (i.e.
+the relation is the set of pairs $(a, b)$ such that $a=b$ or $a\leq b$ respectively).
+This exemplifies the effort in set theory to find general structures.
+\end{remark}
+
 \begin{defn}[Binary relation]
@@
 
+\marginright{~\\
+	\def\setA{ (0,0) ellipse (5 and 4) }
+	\begin{tikzpicture}[scale = 0.3]
+		\draw \setA;
+		\begin{scope}
+			\clip \setA;
+			\draw (-5, 4) .. controls (0, 2) and (0, -2) .. (-5, -4);
+			\draw (2, 4) .. controls (2, 2) and (0, 1) .. (-1.46, 1);
+			\draw (1, 1.9) .. controls (1.5, 1) and (2, -2.5) .. (0.5, -4);
+			\draw (1.49, -1)  .. controls (2, -0.3) and (4, 0.5) ..  (5, 0);
+		\end{scope}
+		\node at (-4, -4.5) {$A$};
+		\node at (-3.2, 0) {$P_1$};
+		\node at (-0.4, 2.6) {$P_2$};
+		\node at (-0.2, -1.2) {$P_3$};
+		\node at (3, -1.7) {$P_4$};
+		\node at (3, 1.5) {$P_5$};
+	\end{tikzpicture} 
+}
 \begin{defn}[Partitions]
@@
 \end{defn}
-
-TODO: diagram
 

### 2021-05-30 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: sections 2 and 3 from Andrea

--- 2021-04-21 Articles/2020-QuantumLogic/Source/quantumlogic.tex
@@
 
-\author{Gabriele Carcassi\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109.
+\author{Andrea Oldofredi\thanks{Section de Philosophie, Universit\'e de Lausanne, Switzerland.
+E-mail: Andrea.Oldofredi@unil.ch} \and Gabriele Carcassi\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109.
 E-mail: carcassi@umich.edu} \and Christine A.
 Aidala\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109.
-E-mail: caidala@umich.edu} \and Andrea Oldofredi\thanks{Section de Philosophie, Universit\'e de Lausanne, Switzerland.
-E-mail: Andrea.Oldofredi@unil.ch}}
+E-mail: caidala@umich.edu}}
 
@@
 Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it s generally recognized that
-the founding text on Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 von Neumann had already stated that from the algebraic structure of quantum theory it is possible to formulate a new propositional calculus, anticipating seminal ideas that would have been fully expressed of his successive work in collaboration with Garrett Birkhoff.}
+the founding text on Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that von Neumann in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 anticipated that from the algebraic structure of quantum theory it would have been possible to formulate a new propositional calculus.
+These ideas would have been fully expressed of his successive collaboration with Garrett Birkhoff.}
 
@@
 In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography.
-Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now widespread among experts as exemplified e.g.\ by \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
+Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now accepted among experts; for details see e.g.\ by \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
 Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory].
@@
 
-Recently one of us asked in another essay whether quantum physics necessarily implies a revision of classical logic (cf.\ \cite{Oldofredi:2020}).
-Alternatively stated, it was asked whether classical logic must be abandoned in the quantum realm or there are quantum theories compatible with a classical propositional calculus.
+Recently one of us faced the question of whether quantum physics necessarily implies a revision of classical logic (cf.\ \cite{Oldofredi:2020}).
+Alternatively stated, it was asked if classical logic must be abandoned in the quantum realm, or there are quantum theories compatible with a classical propositional calculus.
 Analyzing Bohmian Mechanics, a well-known interpretation of QM, it has been showed that quantum physics does not necessarily involves a departure from classical logic.
-More precisely, it has been argued that in the context of BM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow CL---but also the distributivity law holds, contrary to the case of standard QM.
-This fact is a consequence of the clear ontology of the theory and its its theory of measurement.
+More precisely, it has been argued that in the context of this interpretation of QM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow classical rules---but also the distributivity law holds, contrary to the case of standard QM.
+This fact is a consequence of the clear ontology of the theory and its explanation of quantum measurements.
 
@@
 In what follows we will answer this question in the negative.
-More precisely, here we have a three-fold aim: in the first place, we will analyze the main motivations and justifications for the introduction of quantum logic, arguing that they do not entail the failure of CL in the quantum context.
-In the second place we argue that classical and quantum mechanics share a common algebraic structure which can originate a classical propositional calculus for quantum propositions.
-A position dissimilar to the received view about QL.
-Here we will also show that several important statements in quantum mechanics, i.e.\ those concerned with expectation values, are not captured by QL, while they may be taken into account in a classical logical system.
-Finally, we will argue that it is possible to show that classical propositions may obey a quantum logic.
-Clearly, we will also spell out the reasons for which CL seems not to work in the context of QM.
+More precisely, here we have a two-fold aim: in the first place, we will analyze the main motivations and justifications for the introduction of quantum logic arguing that they do not entail the failure of CL in the quantum context.
+In the second place, we show that from the mathematical structures on which QM is built, it is possible to provide a distributive lattice to quantum propositions, contrary to the received view about QL.
+Such a distributive lattice of quantum statements, moreover, will allow us to take into account also assertions concerning expectation values, which are generally excluded in the context of the standard approach to quantum logic.
+Finally, we will argue that it is possible to demonstrate that classical logical propositions can form a non-distributive lattice.
 All these aspects, to our knowledge, are not sufficiently discussed in literature, thus, the present paper can contribute to shed new light on the relation between logic and quantum mechanics.
@@
 
-The essay is structured as follows: Section \ref{QL} provides a brief outline of the standard approach to QL (readers familiar with Birkhoff and von Neumann's proposal can skip this part), while in Section \ref{Motivations} we review the main motivations that have been given in literature to introduce a quantum logical calculus, explaining also why they do not entail the failure of classical logic in QM.
-In Section \ref{Math} we provide a mathematical argument aimed at showing not only that QM is compatible with classical logic---sharing a common logical structure with classical mechanics---but also that important sentences of quantum theory are excluded by standard QL, as stated a few lines above.\ In Section \ref{Discussion} we discuss the main philosophical implications of the preceding sections, and Section \ref{conc} concludes the paper.
+The essay is structured as follows: Section \ref{QL} provides a brief outline of the standard approach to QL (readers familiar with Birkhoff and von Neumann's proposal can skip this part), while in Section \ref{Motivations} we review the main arguments that have been given in literature to introduce a quantum logical calculus.
+Here we explain  that they do not entail the failure of classical logic in QM.
+In Section \ref{Math} we show not only that quantum logical propositions generate a distributive lattice, but also that important sentences of quantum theory---i.e.\ those concerning expectations values---are excluded by standard QL.\ In Section \ref{conc} we discuss the main philosophical implications of the preceding sections and conclude the paper.
 
@@
 In order to present the standard approach to QL, let us say a few words about the relation between classical logic and classical mechanics.
-It is a well known fact that classical propositional logic is equivalent to a Boolean algebra where propositions (atomic and complex) have truth values ``true'' and ``false'' (or 1 and 0), and the main operations among them are conjunction ``$\wedge$'', disjunction ``$\vee$'' and negation ``$\neg$'', the only unary operation.
+It is a well known fact that classical propositional logic is equivalent to a Boolean algebra where propositions (atomic and complex) have truth values ``true'' and ``false'' (or $\top$ and $\bot$), and the main operations among them are conjunction ``$\wedge$'', disjunction ``$\vee$'' and negation ``$\neg$'', the only unary operation.
 From these logical connectives one can introduce secondary operations, as for instance the material implication ``$\longrightarrow$'', the exclusive or ``$\oplus$'' and logical equivalence ``$\longleftrightarrow$''.
@@
 In particular, the algebra of quantum observables is non-commutative and, from a logical perspective, generates a non-distributive orthocomplemented lattice, which is non-Boolean.
-Referring to this, given the axioms of QM Birkhoff and von Neumann showed how the physical content of the theory does not conform to a classical logical structure.
-The principal aim of their founding essay of quantum logic was to ``discover what logical structures one may hope to find in physical theories which, like quantum mechanics, do not conform to classical logic'' (\cite{vonNeumann:1936}, p.\ 823).
+Referring to this, given the axioms of QM, Birkhoff and von Neumann showed how the physical content of the theory does not conform to a classical logical structure.
+The principal aim of their essay on quantum logic was to ``discover what logical structures one may hope to find in physical theories which, like quantum mechanics, do not conform to classical logic'' (\cite{vonNeumann:1936}, p.\ 823).
 The expression quantum logic in this paper must be clarified, since it refers to a quantum propositional calculus in the form of a ``calculus of linear subspaces with respect to \emph{set products}, \emph{linear sums}, and \emph{orthogonal complements}'' which ``resembles the usual calculus of propositions with respect to \emph{and}, \emph{or}, and \emph{not}'' (\emph{ibid}.), where the logical propositions are associated to measurements, tests on quantum systems.\footnote{Cf.\ \cite{Giuntini:2002} and \cite{DallaChiara:2004} for a systematic introduction to various forms of quantum logics, and to \cite{Engesser:2009} for historical and philosophical discussions on the topic.}
@@
 
-The first remarkable difference between QL and CL is the failure of the distributive law.
-It is a well-known fact that such a law holds in the propositional calculus underlying classical mechanics in virtue of the commutative algebra of classical observables.
+The first remarkable difference between QL and CL is the failure of the distributive law; moreover, analyzing the motivations for which \eqref{DL} does not hold in the quantum context it will be relevant to see also other important differences between these logics.
+It is a well-known fact that the propositional calculus underlying classical mechanics is distributive in virtue of the commutative algebra of classical observables.
 Referring to this, \cite{Giuntini:2002} claim that distributivity fails in quantum logic for two main reasons:
@@
 \item the non-commutativity of quantum observables,
-\item the peculiar behavior of the quantum disjunction---which may be true although neither of the member is true in virtue of the linearity of the Schr\"odinger Equation (SE)---the other major deviation from classical logic.
+\item the peculiar behavior of the quantum disjunction---the second major deviation from classical logic---which may be true although neither of the member is true in virtue of the linearity of the Schr\"odinger Equation (SE).
 \end{itemize}
-\noindent Let us then have a closer look to these motivations for the failure of the distributivity law and more generally of CL in the quantum context.
+\noindent Let us then have a closer look to these motivations for the failure of the distributivity law, and more generally of CL.
 
@@
 From a physical perspective this fact entails that if one performs a measurement of $A$ on a quantum system followed by a measurement of $B$, in general one will obtain different results inverting the order of these observations.
-Moreover, in virtue of the non-commutative structure of quantum theory, to measure an operator $C=A+B$ is not typically equivalent to measure $A$ and $B$ independently and then add the respective results, since $A,B$ will be in general incompatible (cf.\ \cite{David:2015}, p.\ 77).
-This algebraic fact of QM constitutes the formal basis to prove the Heisenberg uncertainty relation, a theorem of quantum theory which reflects the operational inability to simultaneously measure the values of incompatible operators with arbitrary precision.
+Moreover, in virtue of the non-commutative algebra of quantum theory, to measure an operator $C=A+B$ is not typically equivalent to measure $A$ and $B$ independently and then add the respective results, since $A,B$ will be in general incompatible (cf.\ \cite{David:2015}, p.\ 77).
+This feature of QM constitutes the formal basis to prove the Heisenberg uncertainty relation, a theorem of quantum theory which reflects the operational inability to simultaneously measure the values of incompatible operators with arbitrary precision.
 Such a result, in turn, is usually interpreted ontologically in the sense that quantum system do not instantiate definite properties in non-measurement situations (cf.\ \cite{Sakurai1994}).
 
-In addition, as we will explain in the remainder of this section, the failure of distributivity in quantum logic is due to another particular feature of this new propositional calculus, namely on the behavior of the quantum disjunction, which is due to the presence of quantum superpositions.
-Indeed, another remarkable difference between QM  and CM is that quantum systems can be in superposition states as a consequence of the linearity of the Schr\"odinger equation---the fundamental dynamical law of QM---which for a single particle reads:
+In addition, as we will explain in the remainder of this section, the failure of distributivity in quantum logic is due to another particular characteristic of this new propositional calculus, namely the behavior of the quantum disjunction, which is different w.r.t.\ to its classical counterpart in virtue of the presence of quantum superpositions.
+Indeed, another remarkable difference between QM  and CM is that quantum systems can be in superposition states as a consequence of the linearity of the Schr\"odinger equation---the fundamental dynamical law of quantum theory---which for a single particle reads:
 \begin{align}
@@
 \begin{align}
-\label{SSE}
-\psi_s=\alpha\psi_{1}+\beta\psi_{2}
+%\label{SSE}
+\psi_s=\alpha\psi_{1}+\beta\psi_{2} \nonumber
 \end{align} 
@@
 Notably, the new superposed state $\psi_s$ is also a consistent representation of the system.
-Referring to this,  \cite{deRonde:2016} explicitly say that in QL ``differently from the case in classical semantics, a quantum disjunction may be true even if neither of it members is true.
-This reflects, for example, the case in which we are dealing with a state such as that of a spin $1/2$ system which is in a linear combination of states up and down'', as in the example we are going to consider in order to explain the failure of distributivity in the quantum context.
+Referring to this,  \cite{deRonde:2016} explicitly say that in QL
+\begin{quote}
+differently from the case in classical semantics, a quantum disjunction may be true even if neither of it members is true.
+This reflects, for example, the case in which we are dealing with a state such as that of a spin $1/2$ system which is in a linear combination of states up and down.
+\end{quote}
+\noindent This is exactly the case that we are going to consider in order to review the usual explanation provided for the failure of the distributivity law in the quantum context.
 
-To claim that in quantum logic conjunction and disjunction are not distributive means that given three propositions $p, q$ and $r$ the logical law
+To claim that in QL conjunction and disjunction are not distributive means that given three propositions $p, q$ and $r$ the logical law
 \begin{align}
@@
 
-\noindent does not hold.\footnote{In the context of quantum logic, distributivity must be replaced by a weaker law: $(p\wedge q)\vee(p\wedge r)\longrightarrow(p\wedge(q\vee r)).$}
-Let us explain this fact with a simple example taken from \cite{Giuntini:2002}.
-Suppose to consider a $1/2-$spin particle, let's say an electron and assume that this paritlce has spin up along the $x$ axis.
-From QM we know that there are only two possible spin states in which a particle can be found in each axis after a spin measurement, namely either in the up or in the down state.
-Furthermore, since spin operators along different axis do not commute, being incompatible observables---$[S_x, S_y]=[S_x, S_z]=[S_y, S_z]\neq0$---they cannot be simultaneously measured.
+\noindent does not hold.\footnote{In the context of quantum logic, it is usually believed that distributivity must be replaced by a weaker law: $(p\wedge q)\vee(p\wedge r)\longrightarrow(p\wedge(q\vee r)).$}
+Let us illustrate this fact with a simple example taken from \cite{Giuntini:2002}.
+Suppose to consider a $1/2-$spin particle, let's say an electron.
+From QM we know that there are only two possible spin states in which a particle can be found along each axis after a spin measurement, namely either in the up or in the down state; for the case at hand we assume that this particle has $x$-spin up.
+Furthermore, since spin operators along different axes do not commute, being incompatible observables---i.e.\ $[S_x, S_y]=[S_x, S_z]=[S_y, S_z]\neq0$---they cannot be simultaneously measured.
 Then, as a consequence of the Heisenberg uncertainty principle, ``both propositions ``$spin_y$ is up'' and ``$spin_y$ is down'' shall be strongly undetermined.
 However, the disjunction ``either $spin_y$ is up or $spin_y$ is down'' must be true'' (\cite{Giuntini:2002}, pp.\ 133-134), and the same holds for the case of spin in the $z$ direction.
-Therefore, if $p$ in \eqref{DL} represents the proposition ``the electron has $x$-spin up'', $q$ the proposition ``the electron has $y$-spin up'' and $r$  the proposition `the electron has $y$-spin down'', it is possible to see that the first part of \eqref{DL} express a true statement---i.e.\ $p\wedge(q\vee r)=1$---whereas $(p\wedge q)$ and $(p\wedge r)$ are both false in virtue of the incompatibility between the spin operators among different axes, implying that the second part of the distributive law asserts a false statement, i.e.\ $(p\wedge q)\vee(p\wedge r)=0$.
-Consequently, this law fails in the context of quantum theory, since we have $p\wedge(q\vee r)=1\longleftrightarrow(p\wedge q)\vee(p\wedge r)=0$.
-\vspace{2mm}
+Therefore, if in the distributivity law \eqref{DL} we suppose that
+\begin{itemize}
+	\item $p$ represents the proposition ``the electron has $x$-spin up'', 
+	\item $q$ the proposition ``the electron has $y$-spin up'',
+	\item $r$  the proposition ``the electron has $y$-spin down'',
+\end{itemize}
+\noindent it is possible to see that the left hand side of the bi-conditional expresses a true statement---i.e.\ $p\wedge(q\vee r)=\top$---, since we assume that $p$ is true and we know that the disjunction $(q\vee r)$ is also true, whereas $(p\wedge q)$ and $(p\wedge r)$ are both false in virtue of the incompatibility between the spin operators among different axes.
+This fact implies that the right hand side of the distributive law asserts a false statement, i.e.\ $(p\wedge q)\vee(p\wedge r)=\bot$.
+Consequently, this law fails in the context of quantum theory, because we have
+\begin{align}
+\label{contradiction}
+p\wedge(q\vee r)=\top\longleftrightarrow(p\wedge q)\vee(p\wedge r)=\bot. 
+\end{align}
 
-In what follows we will stress that these motivations for QL are not sufficient to claim that CL is rejected by quantum theory.
+In what follows, we will argue that these motivations for the introduction of QL are not sufficient to claim that CL is rejected by quantum theory, contrary the received view about quantum logic.
+More precisely, in the remainder of the section we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members are, and (ii) that \eqref{DL} can be retained in QM, since the argument seen above---usually meant to show its failure---contains serious flaws.
+In the next section, furthermore, we will show that QL can be given a distributive lattice of propositions.
 
-\subsection{Disjunction}
+\subsection{The Quantum Disjunction}
 
-Connection between disjunction and superposition.
+As stated above, there seems to be a strict connection between the features of quantum disjunctions and superposition states.
+Indeed, in quantum logic it is commonly accepted that a disjunction can be true although neither of its members are, as for instance in the example just discussed.
+In this case we assumed that the spin of a particle in the $y$ direction was not determined since no measurement of $S_y$ was performed, thereby the electron under consideration was in a superposition of $y$-spin-up and $y$-spin-down states.
+Then, if we introduce the following propositions
+\begin{description}
+	\item $q =$ ``The particle has spin $y^+$''
+	\item $r =$ ``The particle has spin $y^-$'',
+\end{description}
+\noindent quantum logic establishes that even though neither of these statements have definite truth-value (or neither of them is actually true), their disjunction $(q \vee r)$ is true---violating consequently the bivalence principle, the third main difference between CL and QL.
+The explanation of this fact takes into account spin measurements; indeed, when we actually measure the observable $S_y$ on the quantum particle at hand, we will find with certainty only one of the possible outcomes stated above.
+Thus, it must be the case that after a measurement of the spin along the $y$ direction either $q$ or $r$ will become true.
 
-``A relevant feature of $\vee$ is that, differently from the case in classical semantics, a quantum disjunction may be true even if neither of it members is true.
-This reflects, for example, the case in which we are dealing with a state such as that of a spin 1/2 system which is in a linear combination of states up and down.
-Both propositions, ``the state is up'' and ``the state is down,'' may have no definite truth value (the excluded middle principle is violated), but the disjunction ``the state is up or the state is down'' is a tautology.''
+Nonetheless, it is possible provide an argument claiming that a quantum disjunction may be \emph{not always} true when both its elements have indeterminate truth values, undermining this widespread belief.
+Our reasoning starts from some remarks concerning the very definition of quantum logical proposition.
+
+In classical mechanics one generally considers physical systems as objects having well-defined values for inherent, dynamical and relational properties, these attributes moreover do not depend on any observation---meaning that measuring a magnitude on a classical particle will reveal a pre-existing value of the observed quantity.
+Similarly, the logical propositions referring to classical physical systems describe their actual ontic states, i.e.\ these statements regard state of affairs of the systems under consideration which do not depend on any external observer.
+Alternatively stated, since in the classical case observations reveal the pre-existing values of the qualities attributed to the particles independently of any measurement, we can consider classical experimental propositions as statements about the properties constituting the actual ontic state of classical systems.
+Thus, statements about measurements and assertions concerning the ontic state of classical systems can be conflated without any dangerous metaphysical consequence.
+
+However, in the context of QM this move is not allowed, given that properties of quantum systems acquire definite values only in measurement situations, as we already stated above.
+Consequently, one can attribute a definite value for a specific magnitude to a given system only after the measurement of the quantity under consideration, or after having prepared a quantum object in a specific state.\footnote{In operational quantum theory the preparations are a set of instructions that a certain agent has to follow in order to prepare a quantum system in a certain state.} Quantum mechanics, in virtue of the Born's rule, then provides the probability to obtain a certain result among a set of possible outcomes associated with eigenstates and eigenvalues of the measured observable.
+Thus, quantum logical propositions like ``the particles has spin $y^+$'', ``the velocity of the particle is $v$'', ``the position of the particle is $x$'' \emph{etc.}\ do \emph{not} refer to inherent properties of quantum systems---i.e.\ to their ontic state---but rather they concern the interactions taking place between quantum systems and experimental devices either in preparation procedures or measurement situations.
+These kind of propositions must be completely indeterminate, being neither true nor false, with respect to the actual state of quantum systems when they do not interact, in complete agreement with the physical content of QM.
+
+The complex proposition $q \vee r$ is considered true before any spin measurement in virtue of the conditional claim according to which if one performs a measurement of $S_y$, then either $y^+$ or $y^-$ \emph{will be found}.
+Such a statement, then, refers to a classical macroscopic device which, after the performance of a $y$-spin measurement, will indicate with certainty either the outcome $y$-spin-up or $y$-spin-down.\footnote{Clearly we assume that any measurement will yield with certainty only one result.} Nonetheless, neither such quantum disjunction, nor its atomic components tell us something about the inherent property associated with the observable $S_y$ possessed by a quantum particle independently of any measurement of $S_y$.
+
+Referring to this, it is then possible to highlight another peculiar feature of quantum propositions which is relevant for our argument.
+If we consider the sentence $p=$ ``the particle has spin $x^+$'', we know that this statement can be either true of false only after the actual performance of a $x$-spin measurement on a certain particle\footnote{Or after having prepared the system at hand in a definite $x$-spin state.}: this assertion will be true in the case the experimenter will find the $x^+$ outcome and false otherwise.
+Before the performance of this observation, however, it has an indeterminate truth value.
+Nonetheless, we can say that $p$ has a precise meaning even prior to any $x$-spin measurement, since it may be understood as a sentence referring to the expectation value of the $S_x$ operator.
+Hence, in non-measurement situations the propositions $q, r$ above may be similarly regarded as experimentally equivalent to:
+\begin{description}
+	\item $q' =$ ``The expectation value for $S_y$ is 1/2$\hbar$''
+	\item $r' =$ ``The expectation value for $S_y$ is -1/2$\hbar$''.
+\end{description}
+%As a consequence, proposition $q \vee r$ corresponds to the following statement: 
+%\begin{center}
+%``The absolute value of the expectation value for $S_y$ 1/2$\hbar$''.
+%\end{center}
+In this case, it is straightforward to see that the sentence $q' \vee r'$ does not refer to macroscopic devices, but it says something about the possible values of $S_y$.
+If we have a superposition of the $y$-spin-up and $y$-spin-down states---as in the previous example and thereby a disjunction whose elements are $q'$ and $r'$--- then, the expectation value for $S_y$ will be strictly between $-1/2 \hbar$ and $1/2 \hbar$, and therefore $q' \vee r'$, if referred to expectation values will be false as one would expect, given that $S_y$ may take, before the performance of a certain measurement, any value in this interval.
+Therefore, contrary to the received view in quantum logic, it is not always the case that a quantum disjunction is true even though neither of its members are.
+
+This discussion highlighted a crucial ambiguity in quantum propositions like ``the particle has spin up/down along such and such direction'', since on the one hand they may refer to measurements results, and on the other hand they may concern expectation values for some spin observable.
+In order to disambiguate such experimental propositions, one has to take into account whether we are evaluating propositions regarding expectation values or measurement outcomes.
+This conclusion will be helpful in the following subsection where we show how to keep the distributivity law in quantum logic.
+
+\subsection{Keeping Distributivity in Quantum Mechanics}
+
+As we already said, \cite{Giuntini:2002} have argued that the distributivity law fails in quantum mechanics in virtue of (i) the alleged peculiar behavior of the quantum disjunction and (ii) the non-commutativity among spin operators defined along different axes.
+Recalling their example, we introduced the following three propositions
 
 \begin{description}
-	\item $p =$ ``The particle has spin $z^+$''
-	\item $q =$ ``The particle has spin $z^-$''
+    \item $p =$ ``The state of particle is $x^+$''
+    \item $q =$ ``The state of particle is $y^+$''
+    \item $r =$ ``The state of particle is $y^-$''
 \end{description}
 
-Claim: $p \vee q$ is \textbf{always} true, even if $p$ and $q$ may possibly be both false. The idea is that $p \vee q$ includes also all superpositions of $p$ and $q$.
+\noindent assuming that $p$ is true.
+This fact entails two consequences that should be explicitly stated: on the one hand, at a certain time $t$ an experimenter actually performed a measurement of $S_x$ and found the result $x^+$, and on the other hand, no observations of $S_y$ have been made at $t$.
+Moreover, the story goes, if one would measure the spin along the $y$ axis at later time $t'$, one would either get $y^+$ or $y^-$, hence, we say that $q \vee r \equiv \top$.
+Therefore, $p \wedge (q \vee r)$ is true, since both $p$ and $(q \vee r)$ are.
 
-Counterclaim. The propositions $p$ and $q$ are experimentally equivalent to:
-\begin{description}
-	\item $p =$ ``The expectation value for $S_z$ is 1/2$\hbar$''
-	\item $q =$ ``The expectation value for $S_z$ is -1/2$\hbar$''
-\end{description}
-The proposition $p \vee q$ corresponds to the statement ``The absolute value of the expectation value for $S_z$ 1/2$\hbar$''.
-If we have a superposition of $p$ and $q$, the expectation value for $S_z$ will be strictly between $-1/2 \hbar$ and $1/2 \hbar$, and therefore $p \vee q$, when discussing expectation values, will be false as one would expect.
+However, both the conjunctions on the right hand side of the distributivity law are false because $p \wedge q$ and $ p \wedge r$ refer to incompatible observations: QM does not allow a simultaneous measurement of these quantities, therefore, we have $(p \wedge  q) \vee (p \wedge  r) \equiv \bot$.
+As a consequence, the two sides of the bi-conditional expressing the distributive law generate a contradiction, so that one concludes that \eqref{DL} is not valid in quantum logic, and thereby that the lattice of quantum proposition is non-distributive.
 
-\subsection{Distributivity}
+In order to understand the reasons why this example does not show that distributivity fails in quantum theory, it is worth noting that the left hand side of the distributivity law contains propositions referring to different instants of time: indeed, in our propositions $p, q$ and $r$ we are considering the very same electron, and we assume that it has spin $x^+$, meaning that an agent performed at particular time $t$ a measurement of $S_x$ and found $x^+$ as a determinate outcome.
+Moreover, in virtue of the commutation relation $[S_x, S_y]\neq0$, we know that the particle at hand has indeterminate spin along the $y$ axes at $t$, hence, this quantum object is not in a definite eigenstate of $S_y$.
+Nonetheless, the disjunction $(q \vee r)$ is assumed to be true since if we would measure $S_y$ \emph{at a later time} $t'>t$ we would find only one of the possible results contained stated in $q$ and $r$---i.e.\ either spin $y^+$ or $y^-$.
+Taking into account the right hand side of the distributivity law, however, one is evaluating all the propositions $p, q$ and $r$ \emph{at the very same time} $t$: indeed, we ask whether at $t$ the particle has spin $x^+$ and simultaneously either spin $y^+$ or $y^-$.
+As already said, quantum theory forbids a simultaneous measurement of non-commuting quantities, and thus, we obtain the desired contradiction.
+In sum, \eqref{DL} fails because we considered propositions evaluated at different times $t$ and $t'$ in the left hand side of the distributivity law, whereas the propositions on the right hand side are considered uniquely at time $t$.
+Rewriting \eqref{DL} making explicit the temporal dependence of the three propositions, we immediately say that the above example showing the failure of distributivity in quantum logic is flawed:
 
-The distributivity in quantum mechanics is ``weird'' because (1) it can be true even if none of the elements of the disjunction is true due to the linearity of the Schr\"ondinger equation and (2) quantum observables do not commute.
+$$p_t \wedge (q \vee r)_{t'} \longleftrightarrow (p_t \wedge q_t) \vee (p_t\wedge r_t).$$
+
+\noindent It is in general true that $q, r$ will have a definite truth value at time $t'$, since a measurement of $S_y$ is suppose to happen and, therefore, at $t'$ the disjunction $(q \vee r)_{t'}$ will be true.
+The distributivity law, thus, is violated not in virtue of an inherent quantum mechanical motivation or principle, but simply because we evaluate the three assertions at different times---and thereby we have to do with different statements---in the two sides of the bi-conditional.
+
+Now, as in the previous case of the quantum disjunction, it is possible to argue that the distributivity law is not necessary violated in quantum logic if we take carefully into consideration the meaning of quantum propositions in relation to (i) the physical content of QM, and (ii) their temporal order.
+In the first place, let us introduce a temporal parameter $t$ in order to evaluate the above sentences at the very same instant of time:
 
 \begin{description}
-    \item $p =$ ``The particle has spin $x^+$''
-    \item $q =$ ``The particle has spin $y^+$''
-    \item $r =$ ``The particle has spin $y^-$''
+    \item $p_t =$ ``The state of particle at time $t$ is $x^+$''
+    \item $q_t =$ ``The state of particle at time $t$ is $y^+$''
+    \item $r_t =$ ``The state of particle at time $t$ is $y^-$''.
 \end{description}
 
-Classic: $p \wedge (q \vee r)$ if and only if $(p \wedge  q) \vee (p \wedge r)$
+\noindent Also in this case let us assume that $p_t$ is true.
+Then, in virtue of the formalism of quantum theory, we conclude that at time $t$ both $q_t$ and $r_t$ must be false, since the quantum object under consideration is not in a definite eigenstate of $S_y$ at $t$.
+This fact implies that the left hand side of the distributivity law $p_t \wedge (q_t \vee r_t)$ is false, as well as the right hand side $(p_t \wedge  q_t) \vee (p_t \wedge r_t)$ since both elements of the disjunction are false.
+Thus, taking into account a temporal parameter, it is easy to see that both sides of the bi-conditional expressing the distributivity law have the same truth value
 
-Quantum: Assume $p$ is true, which means we measured spin $x$ to be $+$.
-Since if I measured spin $y$ I would either get $+$ or $-$, we have $q \vee r \equiv \top$.
-Therefore $p \wedge (q \vee r)$ is true.
-On the other hand, both terms in the second disjunctions are false: because $p \wedge  q$ refer to incompatible observables, we have $p \wedge  q \equiv \bot$.
-Therefore the $(p \wedge  q) \vee (p \wedge r) \equiv \bot$.
+$$p_t \wedge (q_t \vee r_t) \equiv \bot \equiv (p_t \wedge q_t) \vee (p_t\wedge r_t)$$
 
-\subsubsection{Reframing}
+\noindent  establishing that \eqref{DL} is \emph{not} violated.
 
-2 casi: considerare
+Finally, let us conclude this section by showing that a failure of the distributivity can happen also in the context of classical mechanics if we take into account three propositions evaluated at different temporal instants in the two sides of the law.
+Suppose we have a ball sitting at position $p$ over a hatch at a certain time $t$ (see Figure 1).
+If the hatch opens at a later time $\hat{t}$, the ball drops, bounces around, until it will rest at either position $q$ or $r$ with equal chance.
 
-\begin{description}
-    \item $p =$ ``The state of particle at time $t$ is $x^+$''
-    \item $q =$ ``The state of particle at time $t$ is $y^+$''
-    \item $r =$ ``The state of particle at time $t$ is $y^-$''
-\end{description}
-
-Then all three assertions are incompatible with each other.
-That is, $p \wedge q \equiv q \wedge r \equiv r \wedge p \equiv \bot$.
-Therefore $p \wedge (q \vee r) \equiv \bot \equiv (p \wedge  q) \vee (p \wedge r)$.
-
-\begin{description}
-    \item $p =$ ``The state of particle at time $t$ is $x^+$''
-    \item $q =$ ``The state of particle at time $t$ is $y^+$''
-    \item $r =$ ``The state of particle at time $t$ is $y^-$''
-    \item $\hat{q} =$ ``The state of particle at time $\hat{t}$ is $y^+$''
-    \item $\hat{r} =$ ``The state of particle at time $\hat{t}$ is $y^-$''
-\end{description}
-
-In this case $p \wedge \hat{q} \not\equiv \bot$ and $p \wedge \hat{r} \not\equiv \bot$.
-Therefore the original claim was $p \wedge (\hat{q} \vee \hat{r})$ if and only if $(p \wedge  q) \vee (p \wedge r)$ which is not the classical logic distributive law.
-Note that, $\hat{q} \vee \hat{r} \equiv \top$ under condition that ``We measure spin $y$ at time $\hat{t}$''.
-
-\subsubsection{Classical example}
 \begin{center}
-\includegraphics[scale=.3]{Balldrop.png}
+\includegraphics[scale=.8]{Ball.eps}
 \end{center}
 
-We have a ball sitting at position $p$ over a hatch. If the hatch opens, the ball drops, bounces around, until it will rest at either position $q$ or $r$ with equal chance. Two cases, as before:
+Firstly, let us consider the following statements:
 
@@
     \item $q =$ ``The position of the ball at time $t$ is $q$''
-    \item $r =$ ``The position of the ball at time $t$ is $r$''
+    \item $r =$ ``The position of the ball at time $t$ is $r$''.
 \end{description}
 
-These are three incompatible assertions. Distributivity holds as before.
+\noindent It is straightforward to see that these are three incompatible assertions: since the ball can have only one possible determinate position in space at a given time $t$---in this example it is assumed that it is located at $p$---this object cannot be localized in two different locations, $p$ and $q$ or $p$ and $r$.
+Since both sides of the bi-conditional defining the distributivity law are false, we conclude that \eqref{DL} holds.
+
+Now, if we assume that the hatch is actually opened and that at later time $\hat{t}$, the ball has settled either in $q$ or $r$. Then, we introduce the new sentences:
 
 \begin{description}
-    \item $p =$ ``The position of the ball at time $t$ is $p$''
-    \item $q =$ ``The position of the ball at time $t$ is $q$''
-    \item $r =$ ``The position of the ball at time $t$ is $r$''
     \item $\hat{q} =$ ``The position of the ball at time $\hat{t}$ is $q$''
-    \item $\hat{r} =$ ``The position of the ball at time $\hat{t}$ is $r$''
+    \item $\hat{r} =$ ``The position of the ball at time $\hat{t}$ is $r$''.
 \end{description}
 
-We assume that, after time $t$, the hatch is opened and that at time $\hat{t}$ the ball has settled.
-As before, $\hat{q} \vee \hat{r} \equiv \top$ because the ball must be in one of the two positions after it dropped, therefore, if $p$ is true, the left side of the bi-conditional is true while the right side is false.
+In this case, at time $\hat{t}$ the disjunction $(\hat{q} \vee \hat{r}) \equiv \top$ because the ball must be in one of the two positions after it dropped.
+Hence, taking into account the left hand side of the distributivity law, if we evaluate $p$ at $t$ and $q, r$ at $\hat{t}$ we see that $p_t \wedge (\hat{q} \vee \hat{r}) \equiv \top$, since $p$ is hold true as well as the disjunction $(\hat{q} \vee \hat{r})$.
+However, if we consider the right hand side of \eqref{DL}, evaluating $p, q$ and $r$ at time $t$---as done in the quantum case---we will trivially obtain that $(p \wedge  q) \vee (p \wedge r)\equiv \bot$.
+Thus, a contradiction can be deduced also in classical mechanics if we evaluate sentences at different times in the two sides of the bi-conditional.
+Nonetheless, one would not say that distributivity fails in the classical context, but rather that we have to evaluate propositions taking into account the same temporal instant in both sides of \eqref{DL}.
 
-%\section{Classical and quantum parallel}
-%
-%Classical state: $\rho(x,p)$
-%
-%Quantum state: $\psi(x)$
-%
-%Classical observable $f(x,p)$. Average value $<f> = \int f(x, p) \rho(x, p) dx dp$. Eigenstate $f(x, p) \rho(x, p) = f_0 \rho(x, p)$.
-%
-%Quantum observable $F$. Average value $<F> = \int \psi^\dagger(x) F \psi(x) dx$. Eigenstate $F \psi(x) = f_0 \psi(x)$.
-%
-%In both cases, average value equal to $f_0$ means that the contributions of all parts of the ensemble combine to the value $f_0$.
-Eigenstate $f_0$ means all parts of the ensemble contribute the same value $f_0$.
-%
-%For example, a classical eigenstate of $H = \frac{p^2}{2m} + \frac{1}{2} k x^2$ with eigenvalue $e_0$ is any function $\rho(x,p)$ that is different from zero only on the ellipses with energy $e_0$.
-That is, all elements of the distribution have the same energy.
-%
-%The statement ``the observable $f$ of the system is between $f_0$ and $f_1$'' is therefore unclear.
-It could be taken to mean ``the average of observable $f$ of the system is between $f_0$ and $f_1$'' or ``the observable $f$ for each part of the ensemble is between $f_0$ and $f_1$''.
-In the first case, we are identifying a value from the real line, using the standard topology and the standard Borel algebra.
-In the second case, we are identifying a set of possible functions based on their support.
-%
-%Now suppose that the $U$ and $V$ are two intervals.
-We have ``the average $<f>$ is in $U \cup V$'' is  equivalent to ``the average $<f>$ is in $U$'' $\vee$ ``the average $<f>$ is in $V$''.
-The average value must be in and and only one set.
-We also have ``the support of $\rho$ is $U \cup V$'' is not equivalent to ``the support of $\rho$ is $U$'' $\vee$ ``the support of $\rho$ is $V$''.
-The dis tribution could span both sets at the same time.
-But the lattice generated by such sets is still a $\sigma$-algebra, and therefore follows classical logic.
-%
-%To check/prove: neither algebras is a sub-algebra of the other; both sub-algebras of the norm-induced one.
+In conclusion, contrary to the received view in quantum logic, we have shown that the distributivity law is not violated in quantum context once the correct temporal order of quantum propositions---and consequently their actual meaning---is considered.
 
@@
 After having critically assessed the general motivations to introduce quantum logic as a consequence of the physical content of QM, in this section we present our main argument according to which it is possible to assign a distributive lattice to quantum propositions.
-Remarkably, in what follows we will see that the non-distributivity of the quantum logic lattice does neither originate from non-commuting observables nor from superpositions of quantum mechanics: it is a property of the lattice of subspaces of \emph{any} vector space and of the lattice of subgroups of any group more in general (cf.\ \cite{Davey:2002}).
+Remarkably, in what follows we will see that the non-distributivity of the quantum logic lattice does neither originate from non-commuting observables nor from superpositions of quantum mechanics: it is a property of the lattice of subspaces of \emph{any} vector space, and of the lattice of subgroups of any group more in general (cf.\ \cite{Davey:2002}).
 To this regard, indeed, we also show that taking into account the space of classical probability distributions one can apply a non-distributive lattice to classical propositions.
@@
 Contrary to the widespread opinion, therefore, we will claim not only that quantum physics does not necessarily entails a revision of classical propositional calculus, but also that it is actually possible to show that quantum propositions can form a distributive lattice.
-Furthermore, in agreement with the example of the violation of the distributivity law in the classical context seen in the previous section, in the remainder of the section we will argue also that the propositions of classical mechanics may obey a non-classical logic.
+Furthermore, in agreement with the example of the violation of the distributivity law in the classical context seen a few lines above, in the remainder of the section we will argue also that the propositions of classical mechanics may obey a non-classical type of logic.
 
@@
 
-From a mathematical perspective, the distributive lattice will be provided in both classical and quantum case by a Borel algebra, which is the smallest $\sigma$-algebra on a set $X$ equipped with a topological structure that contains the topology; the non-distributive lattice will be provided by the set of subgroups ordered by inclusion.
+From a mathematical perspective, the distributive lattice will be provided in both classical and quantum case by a Borel algebra, which is the smallest $\sigma$-algebra on a set $X$ equipped with a topological structure that contains the topology; the non-distributive lattice will be provided by the set of closed subgroups ordered by inclusion.
 We will claim that the non-distributive lattice is a strict subset of the distributive one; this result entails that everything that can be described in quantum logic can also be described in classical logic.
@@
 Hence, measurements of cross sections, the most typical type of measurement in the context of particle physics, are not part of the quantum lattice, and neither are measurements of the probabilities themselves.
-Then, to show that one may apply a distributive lattice to quantum propositions can be seen as an important result in order to extend the set of meaningful logical propositions of quantum mechanics.
-To provide an extended quantum logic lies beyond the introductory scope of the present paper and will be the subject for future research; here we limit ourselves to show that such an extension is possible.
+Therefore, showing that one may apply a distributive lattice to quantum propositions can be seen as an important result in order to extend the set of meaningful logical propositions of quantum mechanics.
 
@@
 
-As anticipated above, in this section our arguments are fundamentally based on the notion of $\sigma$-algebra, which is a significant structure in the context of foundational research, given that physically relevant mathematical branches like measure theory (cf.\ \cite{Cohn:2013}) and probability theory (cf.\ \cite{Grimmett:2001}) are built on top of it.
+As anticipated above, in this section our arguments are fundamentally based on the notion of $\sigma$-algebra, which is a significant structure from a foundational perspective, given that physically relevant mathematical branches like measure theory (cf.\ \cite{Cohn:2013}) and probability theory (cf.\ \cite{Grimmett:2001}) are built on top of it.
 Let us then review some of its essential features.
@@
 As we have already stated, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.\footnote{The choice of the square root of the density over the density itself is merely a technical choice.} In this case we have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
-In all normed spaces, the topology is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.
+In all normed spaces, the topology, and therefore the $\sigma$-algebra, is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.
 
@@
 \begin{prop}
-	Let $H$ be a Hilbert space. Let $\Sigma$ be the Borel $\sigma$-algebra induced by its inner product. Let $L(H)$ be the lattice of closed subspaces of $H$. Then $L(H) \subset \Sigma$.
+	Let $\mathcal{H}$ be a Hilbert space.
+Let $\Sigma$ be the Borel $\sigma$-algebra induced by its inner product.
+Let $L(\mathcal{H})$ be the lattice of closed subspaces of $\mathcal{H}$.
+Then $L(\mathcal{H}) \subset \Sigma$.
 \end{prop}
@@
 Proof.
-Let $p \in L(H)$ be a closed sub-space of $H$.
+Let $p \in L(\mathcal{H})$ be a closed sub-space of $\mathcal{H}$.
 Then $p$ is a closed set, the complement of an open set, and therefore a Borel set: $p \in \Sigma$.
-Let $v \in H$ be a vector in the Hilbert space and consider the singleton $U = \{ v \}$.
-This is not a closed sub-space and therefore $U \notin L(H)$.
-\textcolor{red}{As the topology of a Hilbert space is Hausdorff, all singletons are closed and therefore $U \in \Sigma$.
-This means $L(H) \subset \Sigma$.
-}This concludes the proof.
+Let $v \in \mathcal{H}$ be a vector in the Hilbert space and consider the singleton $U = \{ v \}$.
+This is not a closed sub-space and therefore $U \notin L(\mathcal{H})$.
+As the topology of a Hilbert space is Hausdorff, all singletons are closed and therefore $U \in \Sigma$.\footnote{To give more context, the set $U = \{ v \}$ can be written as $U = \{w \; | \; |w - v|^2 = 0\}$ the set of all elements with zero distance from $v$.
+This can be understood as the limit (i.e.\ countable intersection) of a sequence of propositions ``the object is within $\epsilon$ of $v$'' where $\epsilon$ tends to zero.} This means $L(\mathcal{H}) \subset \Sigma$.
+This concludes the proof.
 
@@
 quantum logic) fails to be a Boolean algebra.
-Stone's representation theorem tells us that any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.
-intersection, union and complement).
+Stone's representation theorem tells us that any Boolean algebra can be expressed as an algebra of sets using the standard set operations (i.e.\ intersection, union and complement) (cf.\ \cite{Davey:2002}).
 For the lattice of subspaces, note that the disjunction $p \vee q$ does not correspond to the set union of the subspaces, but the span.
@@
 \begin{prop}
-	Let $H$ be a Hilbert space and $\langle \, , \, \rangle$ be its inner product.
+	Let $\mathcal{H}$ be a Hilbert space and $\langle \, , \, \rangle$ be its inner product.
 Let $X$ be a self-adjoint bounded linear operator.
-Let $F_X : H \to \mathbb{R}$ be the map defined by $F_X(\psi) = \langle \psi , X \psi \rangle$.
+Let $F_X : \mathcal{H} \to \mathbb{R}$ be the map defined by $F_X(\psi) = \langle \psi , X \psi \rangle$.
 Let $U \subseteq \mathbb{R}$ be a Borel set.
@@
 We want to be clear that we are not claiming that the lattice of subspaces of quantum logic is not of interest, or that it is of equal interest in both classical and quantum mechanics.
-Here we have pointed it out that the argument that quantum mechanics \emph{requires} necessarily a radical departure from classical logic is untenable, given that both theories are built on similar structures.
+Here we have pointed out that the argument that quantum mechanics \emph{requires} necessarily a radical departure from classical logic is untenable, given that both theories are built on similar structures.
 We conclude that such claims should be substantially weakened, in the light of the arguments that we proposed in this section.

### 2021-06-01 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: reshuffled sections 2 and 3 to separate standard claims from ours

@@
 
-In order to present the standard approach to QL, let us say a few words about the relation between classical logic and classical mechanics.
-It is a well known fact that classical propositional logic is equivalent to a Boolean algebra where propositions (atomic and complex) have truth values ``true'' and ``false'' (or $\top$ and $\bot$), and the main operations among them are conjunction ``$\wedge$'', disjunction ``$\vee$'' and negation ``$\neg$'', the only unary operation.
+Since part of the arguments in favor of QL stem from a claimed unsuitability of classical logic in the context of quantum mechanics, let us say a few words about classical logic, classical mechanics and their purported relationship.
+It is a well known result that classical propositional logic is equivalent to a Boolean algebra where propositions (atomic and complex) have truth values ``true'' and ``false'' (or $\top$ and $\bot$), and the main operations among them are conjunction ``$\wedge$'', disjunction ``$\vee$'' and negation ``$\neg$'', the only unary operation.
 From these logical connectives one can introduce secondary operations, as for instance the material implication ``$\longrightarrow$'', the exclusive or ``$\oplus$'' and logical equivalence ``$\longleftrightarrow$''.
 Furthermore, it is important for our discussion to underline that the laws of propositional logic include commutativity, associativity, identity and distributivity with respect to $\wedge$ and $\vee$.
+It is also well established that the algebraic structure underlying classical mechanics is commutative, distributive and associative.
+The standard view in QL is that classical logic is applicable to classical mechanics because of the commutative algebraic structure.
 
-Analogously, the algebraic structure underlying classical mechanics is Boolean as well, being commutative, distributive and associative; in this context the logical operations of conjunction, disjunction and negation are replaced respectively by multiplication, addition and complementation among the variables of the theory.
-This fact has an interesting physical significance, since the variables associated to magnitudes of physical systems can be added and multiplied together---i.e.\ one can sum and multiply measurement results.
+The idea is that the logical operations of conjunction, disjunction and negation are replaced respectively by multiplication, addition and complementation among the variables of the theory.
+This has an interesting physical significance, since the variables associated to magnitudes of physical systems can be added and multiplied together---i.e.\ one can sum and multiply measurement results.
 Both propositional logic and classical mechanics, thus, generate a complemented Boolean lattice; from this fact it follows that the observable algebra of CM ``is isomorphic to a Boolean algebra of propositions with $(\wedge, \vee, \neg)$'' (\cite{David:2015}, p.
@@
 
-Taking instead into account the mathematical structure and the physical content of standard QM, it is generally claimed that the classical propositional calculus is no longer appropriate to represent the logic obeyed by quantum propositions, since quantum theory does not share the same algebraic structure of classical mechanics.
-In particular, the algebra of quantum observables is non-commutative and, from a logical perspective, generates a non-distributive orthocomplemented lattice, which is non-Boolean.
-Referring to this, given the axioms of QM, Birkhoff and von Neumann showed how the physical content of the theory does not conform to a classical logical structure.
+Once the previous stance is taken, one necessarily concludes that classical propositional calculus is not appropriate to represent the logic obeyed by quantum propositions, since quantum theory does not share the same algebraic structure of classical mechanics.
+Referring to this, \cite{Giuntini:2002} claim that distributivity, a fundamental property of classical logic, fails in quantum logic for two main reasons:
+\begin{itemize}
+	\item the non-commutativity of quantum observables,
+	\item the peculiar behavior of the quantum disjunction---the second major deviation from classical logic---which may be true even though neither of the members is true in virtue of the linearity of the Schr\"odinger Equation (SE).
+\end{itemize}
+\noindent Let us then have a closer look at these motivations for the failure of the distributivity law, and more generally of CL.
+
+Contrary to the case of classical mechanics, QM relies on a non-commutative algebraic structure\footnote{For details on the mathematical structure of QM and its physical content the reader may refer e.g.\ to \cite{Sakurai1994}, and \cite{Griffiths:2014}.
+In this essay, it is assumed that the reader has some familiarity with the standard formalism of QM.}: quantum observables generally do not commute, meaning that for any pair of operators $A,B$ we have $[A,B]=AB-BA\neq0$.
+From a physical perspective this fact entails that if one performs a measurement of $A$ on a quantum system followed by a measurement of $B$, in general one will obtain different results inverting the order of these observations.
+Moreover, in virtue of the non-commutative algebra of quantum theory, to measure an operator $C=A+B$ is not typically equivalent to measure $A$ and $B$ independently and then add the respective results, since $A,B$ will be in general incompatible (cf.\ \cite{David:2015}, p.\ 77).
+This feature of QM constitutes the formal basis to prove the Heisenberg uncertainty relation, a theorem of quantum theory which reflects the operational inability to simultaneously prepare/measure the values of incompatible operators with arbitrary precision.
+Such a result, in turn, is usually interpreted ontologically in the sense that quantum systems do not instantiate definite properties in non-measurement situations (cf.\ \cite{Sakurai1994}).
+
+In addition, as we will explain in the remainder of this section, the failure of distributivity in quantum logic is due to another particular characteristic of this new propositional calculus, namely the behavior of the quantum disjunction, which is different w.r.t.\ its classical counterpart in virtue of the presence of quantum superpositions.
+Indeed, another remarkable difference between QM  and CM is that quantum systems can be in a superposition of states as a consequence of the linearity of the Schr\"odinger equation---the fundamental dynamical law of quantum theory---which for a single particle reads:
+\begin{align}
+	\label{SE}
+	i\hbar\frac{\partial\psi}{\partial t}=\Big(-\frac{\hbar^2}{2m}\nabla_k^2+V\Big)\psi=H\psi,
+\end{align}
+\noindent where $H$ represents the Hamiltonian operator, defined as the sum of kinetic and potential energy of the system at hand.
+More specifically, this algebraic property of \eqref{SE} entails that if two wave functions\footnote{In QM the wave function of a system provides the maximal information available about it.} $\psi_1, \psi_2$ are both possible solutions of the same Sch\"odinger equation, then their linear combination (\emph{superposition})
+\begin{align}
+	%\label{SSE}
+	\psi_s=\alpha\psi_{1}+\beta\psi_{2} \nonumber
+\end{align} 
+is still a solution of the same SE---$|\alpha|^2,|\beta|^2$ (with $\alpha,\beta\in\mathbb{C}$) represent the probabilities to find the system in $\psi_1, \psi_2$ respectively.
+Notably, the new superposed state $\psi_s$ is also a consistent representation of the system.
+Referring to this,  \cite{deRonde:2016} explicitly say that in QL
+\begin{quote}
+	differently from the case in classical semantics, a quantum disjunction may be true even if neither of its members is true.
+This reflects, for example, the case in which we are dealing with a state such as that of a spin $1/2$ system which is in a linear combination of states up and down.
+\end{quote}
+\noindent Let us use this exact case to review both alleged problems, the peculiar nature of the quantum disjunction and the lack of distributivity in quantum logic, with a simple example taken from \cite{Giuntini:2002}.
+
+Let us consider a $1/2-$spin particle, such as an electron.
+From QM we know that there are only two possible spin states in which a particle can be found along each axis after a spin measurement, namely either in the up or in the down state.
+Furthermore, since spin operators along different axes are incompatible and do not commute---i.e.\ $[S_x, S_y]\neq 0, [S_x, S_z]\neq 0, [S_y, S_z]\neq0$---they represent incompatible observables that cannot be simultaneously measured.
+Let us then consider the following scenario: an electron is prepared with spin along the $x$ direction and subsequently the spin along the $y$ axis is measured.
+Let the propositions $q$, $p$ and $r$ be defined as follows:
+\begin{description}
+	\item $p$: ``the electron has $x$-spin up'', 
+	\item $q$: ``the electron has $y$-spin up'',
+	\item $r$: ``the electron has $y$-spin down''.
+\end{description}
+As we said, there are only two states in which a particle can be found along an axis.
+Therefore the proposition $q \vee r$ ``must be true'' (\cite{Giuntini:2002}, pp.\ 133-134).
+However, because the different directions are incompatible, if $p$ is true then both $q$ and $r$ must be false.
+The result is, as noted before, that the quantum disjunction is true even if neither of its members is true.
+
+This behavior in turn breaks the law of distributivity which would state:
+\begin{align}
+	\label{DL}
+	p\wedge(q\vee r)\longleftrightarrow(p\wedge q)\vee(p\wedge r).
+\end{align}
+\noindent Since we assume that $p$ is true and we know that the disjunction $(q\vee r)$ is also true, whereas $(p\wedge q)$ and $(p\wedge r)$ are both false in virtue of the incompatibility between the spin operators among different axes.
+This fact implies that the right hand side of the distributive law asserts a false statement, i.e.\ $(p\wedge q)\vee(p\wedge r)=\bot$.
+Consequently, we have
+\begin{align}
+	\label{contradiction}
+	p\wedge(q\vee r)=\top\longleftrightarrow(p\wedge q)\vee(p\wedge r)=\bot. 
+\end{align}
+Therefore the distributive law fails in the context of quantum theory.\footnote{In the context of quantum logic, it is usually believed that distributivity must be replaced by a weaker law: $(p\wedge q)\vee(p\wedge r)\longrightarrow(p\wedge(q\vee r)).$}
+
+These types of problems were highlighted by Birkhoff and von Neumann, who took them as evidence for the failure of classical logic in the context of quantum theory.
 The principal aim of their essay on quantum logic was to ``discover what logical structures one may hope to find in physical theories which, like quantum mechanics, do not conform to classical logic'' (\cite{vonNeumann:1936}, p.\ 823).
-The expression quantum logic in this paper must be clarified, since it refers to a quantum propositional calculus in the form of a ``calculus of linear subspaces with respect to \emph{set products}, \emph{linear sums}, and \emph{orthogonal complements}'' which ``resembles the usual calculus of propositions with respect to \emph{and}, \emph{or}, and \emph{not}'' (\emph{ibid}.), where the logical propositions are associated to measurements, tests on quantum systems.\footnote{Cf.\ \cite{Giuntini:2002} and \cite{DallaChiara:2004} for a systematic introduction to various forms of quantum logics, and to \cite{Engesser:2009} for historical and philosophical discussions on the topic.}
+The expression \emph{quantum logic} in this paper must be clarified, since it refers to a quantum propositional calculus in the form of a ``calculus of linear subspaces with respect to \emph{set products}, \emph{linear sums}, and \emph{orthogonal complements}'' which ``resembles the usual calculus of propositions with respect to \emph{and}, \emph{or}, and \emph{not}'' (\emph{ibid}.), where the logical propositions are associated to measurements, tests on quantum systems.\footnote{Cf.\ \cite{Giuntini:2002} and \cite{DallaChiara:2004} for a systematic introduction to various forms of quantum logic, and to \cite{Engesser:2009} for historical and philosophical discussions on the topic.}
 
@@
 A point in phase space corresponds to the specification of the position and momentum variables of a certain classical system, whereas in QM the points of $\mathcal{H}$ correspond to wave functions.
-The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcome in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.} as underlined by Svozi:
+The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcome in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.
+[TODO: check quote as it is grammatically incorrect]} as underlined by Svozi:
 \begin{quote}
@@
 Thus, since in quantum mechanics physical systems do not have well-defined values for their properties before the measurement of a given operator (contrary to their classical counterparts), or better that quantum observations do not reveal pre-existing values, we have to stress that such quantum propositions refer always to measurements or to preparation procedures.
-Hence, the sentence ``the particle $s$ has $x$-spin up'' means either that I have measured the observable $S_x$ and I have found the value $+\hbar/2$ associated with the state ``$x$-spin-up'', or equivalently that I prepared $s$ in the $x$-spin-up state.
+Hence, the sentence ``the particle $s$ has $x$-spin up'' means either that we have measured the observable $S_x$ and we have found the value $+\hbar/2$ associated with the state ``$x$-spin-up'', or equivalently that we prepared $s$ in the $x$-spin-up state.
 Thus, one cannot simply say, as in classical mechanics, that a certain system has a certain property without having prepared such a system in a given state or having measured it.\footnote{More on this below.}
@@
 Thus, quantum logic does not introduce a new logical operator for the conjunction.
-\item Contrary to the previous case, the logical operator for disjunction cannot be represented by the set-theoretic union, since the set-theoretical union of the two subspaces $P, Q$ will not be in general a subspace, thus, it is not an experimental proposition.
+\item Contrary to the previous case, the logical operator for disjunction cannot be represented by the set-theoretic union, since the set-theoretic union of the two subspaces $P, Q$ will not be in general a subspace, thus, it is not an experimental proposition.
 Therefore, in QL one introduces the quantum logical disjunction as the the closed span of the subspaces $P, Q$, which is an experimental proposition.
 Such a statement corresponds to the ``smallest closed subspace containing both $P$ and $Q$'' (\cite{Bacciagaluppi:2009}, p.\ 55).
-\item The logical implication is defined by set theoretical inclusion: given two experimental propositions $p$ and $q$, 
-$p$ implies $q$, means that whenever one predicts $p$ with certainty, one can predict also $q$ with certainty, and this is equivalent to state that $p$ is a subset of $q$.
+\item The logical implication is defined by set theoretic inclusion: given two experimental propositions $p$ and $q$, 
+$p$ implies $q$ means that whenever one predicts $p$ with certainty, one can predict also $q$ with certainty, and this is equivalent to stating that $p$ is a subset of $q$.
 This fact is particularly important since the authors showed that ``it is \emph{algebraically} reasonable to try to correlate physical qualities with subsets of phase-space'' and thus, ``physical qualities attributable to any physical system form a partially ordered system'' (\cite{vonNeumann:1936}, p.\ 828, notation adapted).
@@
 
-In this manner we have defined a propositional calculus for quantum experimental propositions, which generates an orthocomplemented lattice.
-In what follows we analyze the motivations for such a quantum logic, clarifying why it is generally believed that classical propositional calculus cannot be applied in QM.
+We have thus reviewed the standard arguments to support the failure of classical logic in the context of quantum mechanics, and we have reviewed the basic mathematical structure used in quantum logic.
+
+% In the following section we will see that Hilbert spaces already posses a mathematical structure that follows classical logic, namely the Borel $\sigma$-algebra, and therefore quantum mechanics already posses a propositional lattice that follows classical logic.
+Moreover, we show that classical probabilistic theories also feature a non-distributive lattice that is the analogue of quantum logic, which means non-distributivity is neither required nor unique to quantum mechanics.
 
@@
 
-The first remarkable difference between QL and CL is the failure of the distributive law; moreover, analyzing the motivations for which \eqref{DL} does not hold in the quantum context it will be relevant to see also other important differences between these logics.
-It is a well-known fact that the propositional calculus underlying classical mechanics is distributive in virtue of the commutative algebra of classical observables.
-Referring to this, \cite{Giuntini:2002} claim that distributivity fails in quantum logic for two main reasons:
-\begin{itemize}
-\item the non-commutativity of quantum observables,
-\item the peculiar behavior of the quantum disjunction---the second major deviation from classical logic---which may be true although neither of the member is true in virtue of the linearity of the Schr\"odinger Equation (SE).
-\end{itemize}
-\noindent Let us then have a closer look to these motivations for the failure of the distributivity law, and more generally of CL.
-
-Contrary to the case of classical mechanics, QM relies on a non-commutative algebraic structure\footnote{For details on the mathematical structure of QM and its physical content the reader may refer e.g.\ to \cite{Sakurai1994}, and \cite{Griffiths:2014}.
-In this essay, it is assumed that the reader has some familiarity with the standard formalism of QM.}: quantum observables generally do not commute, meaning that for any pair of operators $A,B$ we have $[A,B]=AB-BA\neq0$.
-From a physical perspective this fact entails that if one performs a measurement of $A$ on a quantum system followed by a measurement of $B$, in general one will obtain different results inverting the order of these observations.
-Moreover, in virtue of the non-commutative algebra of quantum theory, to measure an operator $C=A+B$ is not typically equivalent to measure $A$ and $B$ independently and then add the respective results, since $A,B$ will be in general incompatible (cf.\ \cite{David:2015}, p.\ 77).
-This feature of QM constitutes the formal basis to prove the Heisenberg uncertainty relation, a theorem of quantum theory which reflects the operational inability to simultaneously measure the values of incompatible operators with arbitrary precision.
-Such a result, in turn, is usually interpreted ontologically in the sense that quantum system do not instantiate definite properties in non-measurement situations (cf.\ \cite{Sakurai1994}).
-
-In addition, as we will explain in the remainder of this section, the failure of distributivity in quantum logic is due to another particular characteristic of this new propositional calculus, namely the behavior of the quantum disjunction, which is different w.r.t.\ to its classical counterpart in virtue of the presence of quantum superpositions.
-Indeed, another remarkable difference between QM  and CM is that quantum systems can be in superposition states as a consequence of the linearity of the Schr\"odinger equation---the fundamental dynamical law of quantum theory---which for a single particle reads:
-\begin{align}
-\label{SE}
-i\hbar\frac{\partial\psi}{\partial t}=\Big(-\frac{\hbar^2}{2m}\nabla_k^2+V\Big)\psi=H\psi,
-\end{align}
-\noindent where $H$ represents the Hamiltonian operator, defined as the sum of kinetic and potential energy of the system at hand.
-More specifically, this algebraic property of \eqref{SE} entails that if two wave functions\footnote{In QM the wave function of a system provides the maximal information available about it.} $\psi_1, \psi_2$ are both possible solutions of the same Sch\"odinger equation, then their linear combination (\emph{superposition})
-\begin{align}
-%\label{SSE}
-\psi_s=\alpha\psi_{1}+\beta\psi_{2} \nonumber
-\end{align} 
-is still a solution of the same SE---$|\alpha|^2,|\beta|^2$ (with $\alpha,\beta\in\mathbb{C}$) represent the probabilities to find the system in $\psi_1, \psi_2$ respectively.
-Notably, the new superposed state $\psi_s$ is also a consistent representation of the system.
-Referring to this,  \cite{deRonde:2016} explicitly say that in QL
-\begin{quote}
-differently from the case in classical semantics, a quantum disjunction may be true even if neither of it members is true.
-This reflects, for example, the case in which we are dealing with a state such as that of a spin $1/2$ system which is in a linear combination of states up and down.
-\end{quote}
-\noindent This is exactly the case that we are going to consider in order to review the usual explanation provided for the failure of the distributivity law in the quantum context.
-
-To claim that in QL conjunction and disjunction are not distributive means that given three propositions $p, q$ and $r$ the logical law
-\begin{align}
-\label{DL}
-p\wedge(q\vee r)\longleftrightarrow(p\wedge q)\vee(p\wedge r)
-\end{align}
-
-\noindent does not hold.\footnote{In the context of quantum logic, it is usually believed that distributivity must be replaced by a weaker law: $(p\wedge q)\vee(p\wedge r)\longrightarrow(p\wedge(q\vee r)).$}
-Let us illustrate this fact with a simple example taken from \cite{Giuntini:2002}.
-Suppose to consider a $1/2-$spin particle, let's say an electron.
-From QM we know that there are only two possible spin states in which a particle can be found along each axis after a spin measurement, namely either in the up or in the down state; for the case at hand we assume that this particle has $x$-spin up.
-Furthermore, since spin operators along different axes do not commute, being incompatible observables---i.e.\ $[S_x, S_y]=[S_x, S_z]=[S_y, S_z]\neq0$---they cannot be simultaneously measured.
-Then, as a consequence of the Heisenberg uncertainty principle, ``both propositions ``$spin_y$ is up'' and ``$spin_y$ is down'' shall be strongly undetermined.
-However, the disjunction ``either $spin_y$ is up or $spin_y$ is down'' must be true'' (\cite{Giuntini:2002}, pp.\ 133-134), and the same holds for the case of spin in the $z$ direction.
-Therefore, if in the distributivity law \eqref{DL} we suppose that
-\begin{itemize}
-	\item $p$ represents the proposition ``the electron has $x$-spin up'', 
-	\item $q$ the proposition ``the electron has $y$-spin up'',
-	\item $r$  the proposition ``the electron has $y$-spin down'',
-\end{itemize}
-\noindent it is possible to see that the left hand side of the bi-conditional expresses a true statement---i.e.\ $p\wedge(q\vee r)=\top$---, since we assume that $p$ is true and we know that the disjunction $(q\vee r)$ is also true, whereas $(p\wedge q)$ and $(p\wedge r)$ are both false in virtue of the incompatibility between the spin operators among different axes.
-This fact implies that the right hand side of the distributive law asserts a false statement, i.e.\ $(p\wedge q)\vee(p\wedge r)=\bot$.
-Consequently, this law fails in the context of quantum theory, because we have
-\begin{align}
-\label{contradiction}
-p\wedge(q\vee r)=\top\longleftrightarrow(p\wedge q)\vee(p\wedge r)=\bot. 
-\end{align}
-
-In what follows, we will argue that these motivations for the introduction of QL are not sufficient to claim that CL is rejected by quantum theory, contrary the received view about quantum logic.
-More precisely, in the remainder of the section we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members are, and (ii) that \eqref{DL} can be retained in QM, since the argument seen above---usually meant to show its failure---contains serious flaws.
-In the next section, furthermore, we will show that QL can be given a distributive lattice of propositions.
+In this section we will review the examples given in the previous one and see that the failure of distributivity stems from the ambiguity of the propositions and, once that is solved, standard classical rules can be applied.
+More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members is, and (ii) that \eqref{DL} can be retained in QM, since the arguments seen above---usually meant to show its failure---contain serious flaws.
 
@@
 
-As stated above, there seems to be a strict connection between the features of quantum disjunctions and superposition states.
-Indeed, in quantum logic it is commonly accepted that a disjunction can be true although neither of its members are, as for instance in the example just discussed.
-In this case we assumed that the spin of a particle in the $y$ direction was not determined since no measurement of $S_y$ was performed, thereby the electron under consideration was in a superposition of $y$-spin-up and $y$-spin-down states.
-Then, if we introduce the following propositions
+As noted in the previous section, in classical mechanics one generally considers physical systems as objects having well-defined values for inherent, dynamical and relational properties which do not depend on any observation---meaning that measuring a magnitude on a classical particle will reveal a pre-existing value of the observed quantity.
+Thus, statements about the ontic state of classical systems before or after a measurement can be conflated without any dangerous metaphysical consequence.
+
+However, in the context of QM this move is not allowed.
+The effect of the measurement on the state of the system, however one may choose to interpret them, cannot be disregarded.
+A statement about a system before a measurement, about its preparation\footnote{In operational quantum theory the preparations are a set of instructions that a certain agent has to follow in order to prepare a quantum system in a certain state.}, is different from a statement about the same system after its measurement, about the measurement outcome.
+In fact, it is only through repeated measurements that we can infer anything about the state of the system before the measurement.
+Thus, propositions like ``the electron has $y$-spin up'', ``the velocity of the particle is $v$'', ``the position of the particle is $x$'' are ambiguous as they do not spell out whether they are about preparation or measurement outcome.
+Once that is taken into account, the purported failure of classical logic disappears.
+
+Let us then redefine the propositions as follows:
 \begin{description}
-	\item $q =$ ``The particle has spin $y^+$''
-	\item $r =$ ``The particle has spin $y^-$'',
+	\item $p_i$: ``before the measurement, the electron $x$-spin is up'', 
+	\item $p_o$: ``after the measurement, the electron $x$-spin is up'', 
+	\item $q_i$: ``before the measurement, the electron $y$-spin is up'',
+	\item $q_o$: ``after the measurement, the electron $y$-spin is up'',
+	\item $r_i$: ``before the measurement, the electron $y$-spin is down'',
+	\item $r_o$: ``after the measurement, the electron $y$-spin is down''.
 \end{description}
-\noindent quantum logic establishes that even though neither of these statements have definite truth-value (or neither of them is actually true), their disjunction $(q \vee r)$ is true---violating consequently the bivalence principle, the third main difference between CL and QL.
-The explanation of this fact takes into account spin measurements; indeed, when we actually measure the observable $S_y$ on the quantum particle at hand, we will find with certainty only one of the possible outcomes stated above.
-Thus, it must be the case that after a measurement of the spin along the $y$ direction either $q$ or $r$ will become true.
+\noindent Given the setup of the example, after the measurement either $q_o$ or $r_o$ must be true, which means $q_o \vee r_o = \top$.
+This reflects the simple fact that after a measurement of spin along the $y$ axis the state must be one of the two, which is true only after the measurement.
+The key is that $q_i$ and $r_i$ are not the same statements as $q_o$ and $r_o$, therefore it is not necessarily true that $q_o \vee r_o = q_i \vee r_i$.
+This unwarranted generalization is what causes the problem.
 
-Nonetheless, it is possible provide an argument claiming that a quantum disjunction may be \emph{not always} true when both its elements have indeterminate truth values, undermining this widespread belief.
-Our reasoning starts from some remarks concerning the very definition of quantum logical proposition.
+To make evident why the generalization is problematic, let us add a third step.
+After the measurement, if the electron has $y$-spin down we will rotate by 90 degrees using the precession generated by a magnetic field, so that $x$-spin will be up.
+If we measure $y$-spin up, however, we do not do anything.
+Consider now the statements $p_f$, $q_f$ and $r_f$ that assert the same properties as before but after the third step.
+By construction, the electron will either have $x$-spin up or $y$-spin up, either $p_f$ or $q_f$ must be true and therefore $p_f \vee q_f = \top$.
+But again, this relationship is valid only after the third step, which means $p_o \vee q_o$ is not necessarily true.
+We could even increase the rotation to 180 degrees, such that the $y$-spin down after the measurement becomes $y$-spin up.
+In that case we would have $q_f = \top$.
+Clearly, the fact that we can construct a process that results in the condition $q_f$ to be always satisfied does not justify us to assume that the same condition is satisfied in all cases.
+But that is exactly what the previous generalization was doing: since after a measurement the $y$-spin is either up or down, it is always either up or down.
+And this does not follow.
+Neither the disjunction, nor its atomic components tell us anything about an inherent property associated with the observable $S_y$ possessed by a quantum particle independently of any measurement of $S_y$.
 
-In classical mechanics one generally considers physical systems as objects having well-defined values for inherent, dynamical and relational properties, these attributes moreover do not depend on any observation---meaning that measuring a magnitude on a classical particle will reveal a pre-existing value of the observed quantity.
-Similarly, the logical propositions referring to classical physical systems describe their actual ontic states, i.e.\ these statements regard state of affairs of the systems under consideration which do not depend on any external observer.
-Alternatively stated, since in the classical case observations reveal the pre-existing values of the qualities attributed to the particles independently of any measurement, we can consider classical experimental propositions as statements about the properties constituting the actual ontic state of classical systems.
-Thus, statements about measurements and assertions concerning the ontic state of classical systems can be conflated without any dangerous metaphysical consequence.
+On the other hand, we can calculate the truth value of $q_i \vee r_i$ directly and show that it is false.
+The difficulty here is that before the measurement only $x$-spin is defined, so how are we going to compute statements about the value of $y$-spin?
+The solution is that while the value of an observable is not always defined, its expectation value always is.
+The idea is then to translate $q_i$ and $r_i$ into propositions about expectation values, which we already know how to handle, and apply the standard rules.
+Consider the following statements:
+\begin{description}
+	\item $q'_i$: ``before the measurement, the expectation value for $y$-spin is $\frac{1}{2} \hbar$''
+	\item $r'_i$: ``before the measurement, the expectation value for $y$-spin is $-\frac{1}{2} \hbar$''.
+\end{description}
+\noindent Because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true.
+Therefore $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$.
+Consequently $q_i \vee r_i = q'_i \vee r'_i$.
+Note that $q'_i \vee r'_i$  is the proposition ``before the measurement, the expectation value for $y$-spin is $\pm \frac{1}{2} \hbar$''.
+Now, before the measurement we know that the electron $x$-spin is up.
+While the $y$-spin value is not fully defined---the state is not an eigenstate of $S_y$---the expectation value is fully defined and it is zero.
+Zero is different from $\pm \frac{1}{2} \hbar$ and therefore $q_i \vee r_i = \bot$.
+In fact, since any (non-trivial) superposition of $y$-spin up and down will result in an expectation value strictly between $-\frac{1}{2}\hbar$ and $\frac{1}{2}\hbar$, $q_i \vee r_i$ will be true if and only if either $q_i$ or $r_i$ is true.
+Therefore it is not the case that the quantum disjunction can be true even if none of the elements is true.
 
-However, in the context of QM this move is not allowed, given that properties of quantum systems acquire definite values only in measurement situations, as we already stated above.
-Consequently, one can attribute a definite value for a specific magnitude to a given system only after the measurement of the quantity under consideration, or after having prepared a quantum object in a specific state.\footnote{In operational quantum theory the preparations are a set of instructions that a certain agent has to follow in order to prepare a quantum system in a certain state.} Quantum mechanics, in virtue of the Born's rule, then provides the probability to obtain a certain result among a set of possible outcomes associated with eigenstates and eigenvalues of the measured observable.
-Thus, quantum logical propositions like ``the particles has spin $y^+$'', ``the velocity of the particle is $v$'', ``the position of the particle is $x$'' \emph{etc.}\ do \emph{not} refer to inherent properties of quantum systems---i.e.\ to their ontic state---but rather they concern the interactions taking place between quantum systems and experimental devices either in preparation procedures or measurement situations.
-These kind of propositions must be completely indeterminate, being neither true nor false, with respect to the actual state of quantum systems when they do not interact, in complete agreement with the physical content of QM.
-
-The complex proposition $q \vee r$ is considered true before any spin measurement in virtue of the conditional claim according to which if one performs a measurement of $S_y$, then either $y^+$ or $y^-$ \emph{will be found}.
-Such a statement, then, refers to a classical macroscopic device which, after the performance of a $y$-spin measurement, will indicate with certainty either the outcome $y$-spin-up or $y$-spin-down.\footnote{Clearly we assume that any measurement will yield with certainty only one result.} Nonetheless, neither such quantum disjunction, nor its atomic components tell us something about the inherent property associated with the observable $S_y$ possessed by a quantum particle independently of any measurement of $S_y$.
-
-Referring to this, it is then possible to highlight another peculiar feature of quantum propositions which is relevant for our argument.
-If we consider the sentence $p=$ ``the particle has spin $x^+$'', we know that this statement can be either true of false only after the actual performance of a $x$-spin measurement on a certain particle\footnote{Or after having prepared the system at hand in a definite $x$-spin state.}: this assertion will be true in the case the experimenter will find the $x^+$ outcome and false otherwise.
-Before the performance of this observation, however, it has an indeterminate truth value.
-Nonetheless, we can say that $p$ has a precise meaning even prior to any $x$-spin measurement, since it may be understood as a sentence referring to the expectation value of the $S_x$ operator.
-Hence, in non-measurement situations the propositions $q, r$ above may be similarly regarded as experimentally equivalent to:
-\begin{description}
-	\item $q' =$ ``The expectation value for $S_y$ is 1/2$\hbar$''
-	\item $r' =$ ``The expectation value for $S_y$ is -1/2$\hbar$''.
-\end{description}
-%As a consequence, proposition $q \vee r$ corresponds to the following statement: 
-%\begin{center}
-%``The absolute value of the expectation value for $S_y$ 1/2$\hbar$''.
-%\end{center}
-In this case, it is straightforward to see that the sentence $q' \vee r'$ does not refer to macroscopic devices, but it says something about the possible values of $S_y$.
-If we have a superposition of the $y$-spin-up and $y$-spin-down states---as in the previous example and thereby a disjunction whose elements are $q'$ and $r'$--- then, the expectation value for $S_y$ will be strictly between $-1/2 \hbar$ and $1/2 \hbar$, and therefore $q' \vee r'$, if referred to expectation values will be false as one would expect, given that $S_y$ may take, before the performance of a certain measurement, any value in this interval.
-Therefore, contrary to the received view in quantum logic, it is not always the case that a quantum disjunction is true even though neither of its members are.
-
-This discussion highlighted a crucial ambiguity in quantum propositions like ``the particle has spin up/down along such and such direction'', since on the one hand they may refer to measurements results, and on the other hand they may concern expectation values for some spin observable.
-In order to disambiguate such experimental propositions, one has to take into account whether we are evaluating propositions regarding expectation values or measurement outcomes.
-This conclusion will be helpful in the following subsection where we show how to keep the distributivity law in quantum logic.
+While a full discussion goes beyond the scope of this work, we should mention that it is rather odd to claim that one of the reasons QM fails to follow classical logic is the linearity of the SE given that linear systems abound in classical mechanics.
+All wave equations are linear which include both the SE and classical electromagnetism.
+The Liouville equation, the classical analogue to the SE that governs the evolution of probability distributions under classical Hamiltonian mechanics, is also linear.
+One can create superpositions of classical elecromagnetic fields or classical probability distributions in the same way one does of wave functions, and this does not seem to imply an invalidation of classical logic.
 
@@
 
-As we already said, \cite{Giuntini:2002} have argued that the distributivity law fails in quantum mechanics in virtue of (i) the alleged peculiar behavior of the quantum disjunction and (ii) the non-commutativity among spin operators defined along different axes.
-Recalling their example, we introduced the following three propositions
+Let us now turn our attention to distributivity and try to rewrite \eqref{contradiction} by using the refined propositions that include the temporal dependence.
+The proposition $p$ always referred to preparation, therefore it maps to $p_i$ in all cases.
+The disjunction was argued to be true on the basis of what happened after the measurement, therefore on the left side of the relationship we should use $q_o$ and $r_o$.
+On the right side, it was argued that $p \wedge q$ was false because of the incompatibility between the spin operators among different axis.
+This means we are considering statements at equal time with $p_i$, and therefore we should use $q_i$ and $r_i$.
+We have
+\begin{align}
+	\label{solved_contradiction}
+	p_i \wedge (q_o \vee r_o) = \top \longleftrightarrow (p_i \wedge q_i) \vee (p_i\wedge r_i) = \bot.
+\end{align}
+\noindent But this is hardly a failure of distributivity, since the statements on the left side are not the same as the statements on the right side. Let us fix that.
 
-\begin{description}
-    \item $p =$ ``The state of particle is $x^+$''
-    \item $q =$ ``The state of particle is $y^+$''
-    \item $r =$ ``The state of particle is $y^-$''
-\end{description}
+Before we saw that, if $p_i$ is true, the expectation for $y$-spin is zero, and therefore $q'_i \vee r'_i = q_i \vee r_i = \bot$. Therefore
+\begin{align}
+	\label{solved_distrib_1}
+	p_i \wedge (q_i \vee r_i) = \bot \longleftrightarrow (p_i \wedge q_i) \vee (p_i\wedge r_i) = \bot.
+\end{align}
+\noindent On the other hand, consider $p_i \wedge q_o$ which states that we find $y$-spin up after the measurement and $x$-spin up before.
+This is not false, and in fact it will be true in 50\% of the cases according to the Born rule.
+Similarly, $p_i \wedge r_o$ will be true in the remaining 50\%.
+Therefore we have
+\begin{align}
+	\label{solved_distrib_2}
+	p_i \wedge (q_o \vee r_o) = \top \longleftrightarrow (p_i \wedge q_o) \vee (p_i\wedge r_o) = \top.
+\end{align}
+\noindent In either case, the distributivity law is satisfied.
 
-\noindent assuming that $p$ is true.
-This fact entails two consequences that should be explicitly stated: on the one hand, at a certain time $t$ an experimenter actually performed a measurement of $S_x$ and found the result $x^+$, and on the other hand, no observations of $S_y$ have been made at $t$.
-Moreover, the story goes, if one would measure the spin along the $y$ axis at later time $t'$, one would either get $y^+$ or $y^-$, hence, we say that $q \vee r \equiv \top$.
-Therefore, $p \wedge (q \vee r)$ is true, since both $p$ and $(q \vee r)$ are.
+To sum up, \eqref{DL} fails because we considered propositions evaluated at different times, before and after the measurement.
+If we remove the ambiguity and make the temporal dependence explicit, the distributivity law is recovered.
+The lack of commutativity of the observables in quantum mechanics therefore does not change the rules of logic.
+It imposes that the lattice of propositions at different times depends on the process, in particular on the choice of measurement.
+That is, if at time $t$ a particular observable is either prepared or measured, all statements about incompatible observables will be impossible and therefore they will be false.
+Quantum measurements do have an impact on the lattice of statements, unlike classical measurements, but that impact does not change the rules of logic.
 
-However, both the conjunctions on the right hand side of the distributivity law are false because $p \wedge q$ and $ p \wedge r$ refer to incompatible observations: QM does not allow a simultaneous measurement of these quantities, therefore, we have $(p \wedge  q) \vee (p \wedge  r) \equiv \bot$.
-As a consequence, the two sides of the bi-conditional expressing the distributive law generate a contradiction, so that one concludes that \eqref{DL} is not valid in quantum logic, and thereby that the lattice of quantum proposition is non-distributive.
-
-In order to understand the reasons why this example does not show that distributivity fails in quantum theory, it is worth noting that the left hand side of the distributivity law contains propositions referring to different instants of time: indeed, in our propositions $p, q$ and $r$ we are considering the very same electron, and we assume that it has spin $x^+$, meaning that an agent performed at particular time $t$ a measurement of $S_x$ and found $x^+$ as a determinate outcome.
-Moreover, in virtue of the commutation relation $[S_x, S_y]\neq0$, we know that the particle at hand has indeterminate spin along the $y$ axes at $t$, hence, this quantum object is not in a definite eigenstate of $S_y$.
-Nonetheless, the disjunction $(q \vee r)$ is assumed to be true since if we would measure $S_y$ \emph{at a later time} $t'>t$ we would find only one of the possible results contained stated in $q$ and $r$---i.e.\ either spin $y^+$ or $y^-$.
-Taking into account the right hand side of the distributivity law, however, one is evaluating all the propositions $p, q$ and $r$ \emph{at the very same time} $t$: indeed, we ask whether at $t$ the particle has spin $x^+$ and simultaneously either spin $y^+$ or $y^-$.
-As already said, quantum theory forbids a simultaneous measurement of non-commuting quantities, and thus, we obtain the desired contradiction.
-In sum, \eqref{DL} fails because we considered propositions evaluated at different times $t$ and $t'$ in the left hand side of the distributivity law, whereas the propositions on the right hand side are considered uniquely at time $t$.
-Rewriting \eqref{DL} making explicit the temporal dependence of the three propositions, we immediately say that the above example showing the failure of distributivity in quantum logic is flawed:
-
-$$p_t \wedge (q \vee r)_{t'} \longleftrightarrow (p_t \wedge q_t) \vee (p_t\wedge r_t).$$
-
-\noindent It is in general true that $q, r$ will have a definite truth value at time $t'$, since a measurement of $S_y$ is suppose to happen and, therefore, at $t'$ the disjunction $(q \vee r)_{t'}$ will be true.
-The distributivity law, thus, is violated not in virtue of an inherent quantum mechanical motivation or principle, but simply because we evaluate the three assertions at different times---and thereby we have to do with different statements---in the two sides of the bi-conditional.
-
-Now, as in the previous case of the quantum disjunction, it is possible to argue that the distributivity law is not necessary violated in quantum logic if we take carefully into consideration the meaning of quantum propositions in relation to (i) the physical content of QM, and (ii) their temporal order.
-In the first place, let us introduce a temporal parameter $t$ in order to evaluate the above sentences at the very same instant of time:
-
-\begin{description}
-    \item $p_t =$ ``The state of particle at time $t$ is $x^+$''
-    \item $q_t =$ ``The state of particle at time $t$ is $y^+$''
-    \item $r_t =$ ``The state of particle at time $t$ is $y^-$''.
-\end{description}
-
-\noindent Also in this case let us assume that $p_t$ is true.
-Then, in virtue of the formalism of quantum theory, we conclude that at time $t$ both $q_t$ and $r_t$ must be false, since the quantum object under consideration is not in a definite eigenstate of $S_y$ at $t$.
-This fact implies that the left hand side of the distributivity law $p_t \wedge (q_t \vee r_t)$ is false, as well as the right hand side $(p_t \wedge  q_t) \vee (p_t \wedge r_t)$ since both elements of the disjunction are false.
-Thus, taking into account a temporal parameter, it is easy to see that both sides of the bi-conditional expressing the distributivity law have the same truth value
-
-$$p_t \wedge (q_t \vee r_t) \equiv \bot \equiv (p_t \wedge q_t) \vee (p_t\wedge r_t)$$
-
-\noindent  establishing that \eqref{DL} is \emph{not} violated.
-
-Finally, let us conclude this section by showing that a failure of the distributivity can happen also in the context of classical mechanics if we take into account three propositions evaluated at different temporal instants in the two sides of the law.
-Suppose we have a ball sitting at position $p$ over a hatch at a certain time $t$ (see Figure 1).
-If the hatch opens at a later time $\hat{t}$, the ball drops, bounces around, until it will rest at either position $q$ or $r$ with equal chance.
+The idea that the same statement may or may not be possible at different moments in time is not particular to quantum mechanics.
+We can, in fact, construct a scenario in classical mechanics where propositions have the same before/after logical relationships.
+Suppose we have a ball sitting at position $p$ over a hatch (see Figure 1).
+Once the hatch opens, the ball drops, bounces around, until it rests at either position $q$ or $r$ with equal chance.
 
@@
 
-Firstly, let us consider the following statements:
+Let us consider the following statements:
+\begin{description}
+    \item $p_i$: ``Before the hatch is opened, the position of the ball is $p$''
+    \item $q_i$: ``Before the hatch is opened, the position of the ball is $q$''
+    \item $r_i$: ``Before the hatch is opened, the position of the ball is $r$''.
+\end{description}
+\noindent It is straightforward to see that these are three incompatible assertions: since the ball can have only one possible determinate position in space before the hatch is open, it cannot be localized in two different locations, $p$ and $q$ or $p$ and $r$.
+Therefore the statements obey \eqref{solved_distrib_1}.
+After the hatch is opened, the ball will either settle in $q$ or $r$.
+Consider, then, the sentences:
+\begin{description}
+	\item $p_f$: ``After the hatch is opened, the position of the ball is $p$''
+	\item $q_f$: ``After the hatch is opened, the position of the ball is $q$''
+	\item $r_f$: ``After the hatch is opened, the position of the ball is $r$''.
+\end{description}
+Since the ball must be in one of the two lower positions after it dropped, the disjunction $(q_f \vee r_f) \equiv \top$.
+Therefore the statements obey \eqref{solved_distrib_2}.
+To no surprise, the distributive law is satisfied and therefore classical logic applies.
 
-\begin{description}
-    \item $p =$ ``The position of the ball at time $t$ is $p$''
-    \item $q =$ ``The position of the ball at time $t$ is $q$''
-    \item $r =$ ``The position of the ball at time $t$ is $r$''.
-\end{description}
+We can, however, evaluate $q$ and $r$ at different times, as it is done in the quantum case.
+In that case, we obtain the same relationship as in \eqref{solved_contradiction}.
+If we omit the indices, that is we are vague about the temporal evaluation, we get the same failure of distributivity as in \eqref{contradiction}.
+Naturally, one would not say that distributivity fails in the classical context, but rather that we have to evaluate propositions taking into account the same temporal instant in both sides of \eqref{DL}.
 
-\noindent It is straightforward to see that these are three incompatible assertions: since the ball can have only one possible determinate position in space at a given time $t$---in this example it is assumed that it is located at $p$---this object cannot be localized in two different locations, $p$ and $q$ or $p$ and $r$.
-Since both sides of the bi-conditional defining the distributivity law are false, we conclude that \eqref{DL} holds.
-
-Now, if we assume that the hatch is actually opened and that at later time $\hat{t}$, the ball has settled either in $q$ or $r$. Then, we introduce the new sentences:
-
-\begin{description}
-    \item $\hat{q} =$ ``The position of the ball at time $\hat{t}$ is $q$''
-    \item $\hat{r} =$ ``The position of the ball at time $\hat{t}$ is $r$''.
-\end{description}
-
-In this case, at time $\hat{t}$ the disjunction $(\hat{q} \vee \hat{r}) \equiv \top$ because the ball must be in one of the two positions after it dropped.
-Hence, taking into account the left hand side of the distributivity law, if we evaluate $p$ at $t$ and $q, r$ at $\hat{t}$ we see that $p_t \wedge (\hat{q} \vee \hat{r}) \equiv \top$, since $p$ is hold true as well as the disjunction $(\hat{q} \vee \hat{r})$.
-However, if we consider the right hand side of \eqref{DL}, evaluating $p, q$ and $r$ at time $t$---as done in the quantum case---we will trivially obtain that $(p \wedge  q) \vee (p \wedge r)\equiv \bot$.
-Thus, a contradiction can be deduced also in classical mechanics if we evaluate sentences at different times in the two sides of the bi-conditional.
-Nonetheless, one would not say that distributivity fails in the classical context, but rather that we have to evaluate propositions taking into account the same temporal instant in both sides of \eqref{DL}.
-
-In conclusion, contrary to the received view in quantum logic, we have shown that the distributivity law is not violated in quantum context once the correct temporal order of quantum propositions---and consequently their actual meaning---is considered.
+In conclusion, contrary to the received view in quantum logic, we have shown that the distributivity law is not violated in the quantum context once the correct temporal order of quantum propositions---and consequently their actual meaning---is considered.
 

### 2021-06-10 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: reviewed new example

@@
 To make evident why the generalization is problematic, let us add a third step.
-After the measurement, if the electron has $y$-spin down we will rotate by 90 degrees using the precession generated by a magnetic field, so that $x$-spin will be up.
-If we measure $y$-spin up, however, we do not do anything.
-Consider now the statements $p_f$, $q_f$ and $r_f$ that assert the same properties as before but after the third step.
-By construction, the electron will either have $x$-spin up or $y$-spin up, either $p_f$ or $q_f$ must be true and therefore $p_f \vee q_f = \top$.
-But again, this relationship is valid only after the third step, which means $p_o \vee q_o$ is not necessarily true.
-We could even increase the rotation to 180 degrees, such that the $y$-spin down after the measurement becomes $y$-spin up.
-In that case we would have $q_f = \top$.
-Clearly, the fact that we can construct a process that results in the condition $q_f$ to be always satisfied does not justify us to assume that the same condition is satisfied in all cases.
-But that is exactly what the previous generalization was doing: since after a measurement the $y$-spin is either up or down, it is always either up or down.
-And this does not follow.
+After the measurement of $S_y$, in the case we obtained $y$-spin down we will rotate the direction of spin by 90 degrees using the precession generated by a magnetic field, so that $x$-spin will be up.
+However, in the other case, when $y$-spin up is obtained, we leave the direction of spin unchanged.
+Consider now the statements:
+\begin{description}
+	\item $p_f$: ``after the third step, the electron $x$-spin is up'', 
+	\item $q_f$: ``after the third step, the electron $y$-spin is up'',
+	\item $r_f$: ``after the third step, the electron $y$-spin is down''.
+\end{description}
+\noindent that assert the same properties as before but at the later time.
+By construction, the electron will either have $x$-spin up or $y$-spin up, either $p_f$ or $q_f$ must be true and therefore $p_f \vee q_f = \top$.\footnote{In fact, note that $q_f$ is true if and only if $q_o$ is true, and, similarly, $p_f$ is true if and only if $r_o$ is true.} But again, this relationship is valid only after the third step, which means $p_o \vee q_o$ is not necessarily true because these statements are evaluated at different times.
+But this situation is formally identical to the previous one: before we had that $q_o \vee r_o \neq q_i \vee r_i$, now we have $p_f \vee q_f \neq p_o \vee q_o$.
+It would be improper to say that an electron has always either $x$-spin up or $y$-spin up just because it is so after the third step.
+In the same way, it would be improper to say that the electron as always either $y$-spin up or down just because it is so after the measurement of spin along the $y$ axis.
 Neither the disjunction, nor its atomic components tell us anything about an inherent property associated with the observable $S_y$ possessed by a quantum particle independently of any measurement of $S_y$.

### 2021-06-11 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: added drawing

@@
 \usepackage{natbib}
+
+\usepackage{tikz}
+\usetikzlibrary{calc,positioning,arrows,arrows.meta,decorations.markings}
+
 \usepackage{amsmath, amsthm, amssymb}
@@
 In their view, a typical example of invalid conjunction is one attributing a definite value for both position and momentum to a single quantum particle at a given time $t$.
-Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it s generally recognized that
+Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it is generally recognized that
 the founding text on Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that von Neumann in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 anticipated that from the algebraic structure of quantum theory it would have been possible to formulate a new propositional calculus.
@@
 This unwarranted generalization is what causes the problem.
+
+\tikzstyle{vecArrow} = [thick, decoration={markings,mark=at position
+	1 with {\arrow[semithick]{open triangle 60}}},
+double distance=1.4pt, shorten >= 5.5pt,
+preaction = {decorate},
+postaction = {draw,line width=1.4pt, white,shorten >= 4.5pt}]
+\tikzstyle{innerWhite} = [semithick, white,line width=1.4pt, shorten >= 4.5pt]
+
+\begin{center}
+\scalebox{0.9}{
+	\begin{tikzpicture}
+		\coordinate (o) at (-3,0);
+		\draw[-{Stealth[scale=1.8]}] ($ (o) + (0, -3) $) -- ($ (o) + (0, 3) $);
+		\node at ($ (o) + (.25, 2.5) $) {$y$};
+		\draw[-{Stealth[scale=1.5]}] ($ (o) + (-.985, -.3) $) -- ($ (o) + (.985, .3) $);
+		\node at ($ (o) + (.7, -.2) $) {$x$};
+
+		\coordinate (a) at (0,0);
+		\draw[dashed] ($ (a) + (0, 0) $) ellipse (1 and 3);
+		\draw[-{Stealth[scale=1.8]}, dashed] ($ (a) + (0, 0) $) -- ($ (a) + (0, 3) $);
+		\node at ($ (a) + (-.2, 1.3) $) {$q_i$};
+		\draw[-{Stealth[scale=1.5]}, thick] ($ (a) + (0, 0) $) -- ($ (a) + (.985, .3) $);
+		\node at ($ (a) + (.6, -.1) $) {$p_i$};
+		\draw[-{Stealth[scale=1.8]}, dashed] ($ (a) + (0, 0) $) -- ($ (a) + (0, -3) $);
+		\node at ($ (a) + (-.2, -1.3) $) {$r_i$};
+		\coordinate (aT) at ($ (a) + (0, -3.5) $);
+		\node[align=center] at (aT) {\small Step 1 \\ Preparation};
+
+		\coordinate (bb) at (1.5,0);
+		\draw[vecArrow] ($ (bb) + (-0.2, 0) $) to ($ (bb) + (0.2, 0) $);
+		\draw[innerWhite] ($ (bb) + (-0.2, 0) $) to ($ (bb) + (0.2, 0) $);
+
+		\coordinate (bc) at (3, 0);
+		\draw ($ (bc) + (-1.2, -.6) $) rectangle ($ (bc) + (+1.2, .6) $);
+		\node at ($ (bc) $) {Measurement};
+
+		\coordinate (bd) at (4.5,0);
+		\draw[vecArrow] ($ (bd) + (-0.2, 0) $) to ($ (bd) + (0.2, 0) $);
+		\draw[innerWhite] ($ (bd) + (-0.2, 0) $) to ($ (bd) + (0.2, 0) $);
+
+		\coordinate (b) at (6,0);
+		\draw[dashed] ($ (b) + (0, 0) $) ellipse (1 and 3);
+		\draw[-{Stealth[scale=1.5]}, thick] ($ (b) + (0, 0) $) -- ($ (b) + (0, 3) $);
+		\node at ($ (b) + (-.2, 1.3) $) {$q_o$};
+		\draw[-{Stealth[scale=1.8]}, dashed] ($ (b) + (0, 0) $) -- ($ (b) + (.985, .3) $);
+		\node at ($ (b) + (.6, -.1) $) {$p_o$};
+		\draw[-{Stealth[scale=1.5]}, thick] ($ (b) + (0, 0) $) -- ($ (b) + (0, -3) $);
+		\node at ($ (b) + (-.2, -1.3) $) {$r_o$};
+		\coordinate (bT) at ($ (b) + (0, -3.5) $);
+		\node[align=center] at (bT) {\small Step 2 \\ Measurement};
+
+		\coordinate (cb) at (7.5,0);
+		\draw[vecArrow] ($ (cb) + (-0.2, 1.5) $) to ($ (cb) + (0.2, 1.5) $);
+		\draw[innerWhite] ($ (cb) + (-0.2, 1.5) $) to ($ (cb) + (0.2, 1.5) $);
+		\draw[vecArrow] ($ (cb) + (-0.2, -1.5) $) to ($ (cb) + (0.2, -1.5) $);
+		\draw[innerWhite] ($ (cb) + (-0.2, -1.5) $) to ($ (cb) + (0.2, -1.5) $);
+
+		\coordinate (cc) at (9, 0);
+		\draw[dashed] ($ (cc) + (-1.2, -.6) + (0, 1.5) $) rectangle ($ (cc) + (+1.2, .6) + (0, 1.5) $);
+		\node at ($ (cc) + (0, 1.5) $) {No rotation};
+		\draw ($ (cc) + (-1.2, -.6) + (0, -1.5) $) rectangle ($ (cc) + (+1.2, .6) + (0, -1.5) $);
+		\node at ($ (cc) + (0, -1.5) $) {90$^o$ rotation};
+
+		\coordinate (cd) at (10.5,0);
+		\draw[vecArrow] ($ (cd) + (-0.2, 1.5) $) to ($ (cd) + (0.2, 1.5) $);
+		\draw[innerWhite] ($ (cd) + (-0.2, 1.5) $) to ($ (cd) + (0.2, 1.5) $);
+		\draw[vecArrow] ($ (cd) + (-0.2, -1.5) $) to ($ (cd) + (0.2, -1.5) $);
+		\draw[innerWhite] ($ (cd) + (-0.2, -1.5) $) to ($ (cd) + (0.2, -1.5) $);
+
+		\coordinate (c) at (12,0);
+		\draw[dashed] ($ (c) + (0, 0) $) ellipse (1 and 3);
+		\draw[-{Stealth[scale=1.5]}, thick] ($ (c) + (0, 0) $) -- ($ (c) + (0, 3) $);
+		\node at ($ (c) + (-.2, 1.3) $) {$q_f$};
+		\draw[-{Stealth[scale=1.5]}, thick] ($ (c) + (0, 0) $) -- ($ (c) + (.985, .3) $);
+		\node at ($ (c) + (.6, -.1) $) {$p_f$};
+		\draw[-{Stealth[scale=1.8]}, dashed] ($ (c) + (0, 0) $) -- ($ (c) + (0, -3) $);
+		\node at ($ (c) + (-.2, -1.3) $) {$r_f$};
+		\coordinate (cT) at ($ (c) + (0, -3.5) $);
+		\node[align=center] at (cT) {\small Step 3 \\ Rotation};
+
+		%	\draw \setA;
+		%	\begin{scope}
+		%		\clip \setA;
+		%		\draw (-5, 4) .. controls (0, 2) and (0, -2) .. (-5, -4);
+		%		\draw (2, 4) .. controls (2, 2) and (0, 1) .. (-1.46, 1);
+		%		\draw (1, 1.9) .. controls (1.5, 1) and (2, -2.5) .. (0.5, -4);
+		%		\draw (1.49, -1)  .. controls (2, -0.3) and (4, 0.5) ..  (5, 0);
+		%	\end{scope}
+		%	\node at (-4, -4.5) {$A$};
+		%	\node at (-3.2, 0) {$P_1$};
+		%	\node at (-0.4, 2.6) {$P_2$};
+		%	\node at (-0.2, -1.2) {$P_3$};
+		%	\node at (3, -1.7) {$P_4$};
+		%	\node at (3, 1.5) {$P_5$};
+	\end{tikzpicture} 
+}
+\end{center}
 

### 2021-06-16 Overviews/Process entropy/ProcessEntropy.tex
ProcEntropy: another pass

@@
 it is an ensemble property)?
-What is entropy maximization and why is so successful in many fields, not just physics?
 Is there a general notion of equilibrium that includes both dynamical and thermodynamical equilibria in a way that is relativistically invariant?
+What is entropy maximization, why is so successful in many fields and how is it linked to state definition?
 The working hypothesis is that entropy is the logarithm of the count of the possible evolutions of a system within a specific process.
@@
 This switch from counting states to counting evolutions makes the concept more relativistically sound and allows to properly keep track of how the information about the system is transported back and forward in time by the process, which is essential to apply the concept in a more general setting.
+Defining states requires, at least in some setting, the independence of the system, which indirectly leads to entropy maximization.
 The objective is to formulate a precise mathematical framework that makes this intuition precise and recovers the usual entropies in specific settings.
@@
 Previous work focused on showing how requiring experimental verifiability of a theory leads to topological and $\sigma$-algebraic structures within that theory.
-We have also seen how the ability to compare the level of granularity within the theory is connected to geometric and measure theoretic structures, giving us a way to quantify the information contained in each statement of the theory.
-We now want to characterize the relationship between the granularity of the descriptions at different times (i.e their information content.
+Another effort is exploring how the ability to compare the level of granularity within the theory is connected to geometric and measure theoretic structures, giving us a way to quantify the information contained in each statement of the theory.
+We now want to characterize the relationship between the granularity  (i.e the information content) of the descriptions at different times.
 \textbf{The working hypothesis is that entropy captures how coarse or fine grained is the description of a particular system within a particular process.
-It does so by measuring the logarithm of the count of the possible evolutions that are compatible.} Understanding what are the appropriate basic definitions, how do they map to the standard one and how they are required for the very definitions of systems and states is therefore the central problem of this paper.
+It does so by measuring the logarithm of the count of the possible evolutions that are compatible.} One way to look at it is that as the Boltzmann entropy fails to take into account interactions and correlations between molecules, the count of the microstates and the Shannon/Gibbs entropy fail to take into account system/environment correlations.
+Since these correlations may change in time, the same description of the same system will corresponds to different entropy in time, and this is what the process entropy is designed to solve.
+The central problem of this paper is therefore to understand what are the appropriate basic definitions, how do they map to the standard ones and how system independence, as characterized by entropy, is required for the very definitions of systems and states.
 
-To give a sense of how this would work, we start with defining a \textbf{process domain} $\tdomain[P]$.
-This is a particular instance of our notion of a theoretical domain, and it will represent all meaningful statements, all descriptions at all times and at all scales, that apply to a particular process.
+To give a sense of how this would work, we start with defining a \textbf{process domain} $\tdomain[P]$, which represents all meaningful statements, all descriptions at all times and at all scales, that apply to a particular process.
 Mathematically, it is a $\sigma$-complete Boolean algebra, which means it is closed under negation, countable disjunction (i.e.
@@
 In these cases, the first statement may not be contained or overlap with the second.
-We will assume fineness have all the properties it needs such that if a statement $\stmt[u] \in \tdomain[P]$ is chose, we can construct a measure $\mu_{\stmt[u]}$ that quantifies the granularity of a statement in terms of the unit.
+We will assume fineness have all the properties it needs such that if a statement $\stmt[u] \in \tdomain[P]$ is chosen, we can construct a measure $\mu_{\stmt[u]}$ that quantifies the granularity of a statement in terms of the unit.
 For example, if $\stmt[u]=$ \statement{the position of the ball is within 0 and 1 meters} and $\stmt=$ \statement{the position of the ball is within 3.5 and 5 meters}, then $\mu_{\stmt[u]}(\stmt) = 1.5$.
@@
 This relationship is symmetric with time.
-Similarly we can study how states of a particular system evolve in time, define determinism and reversibility in the terms of the ability within a process to predict or reconstruct states.
+
+Similarly, we can define a trajectory $x(t)$ of a particular system, where $x$ is the full description of that system.
+A process is deterministic if knowing $x(t)$ means we can predict $x(t + \Delta t)$, meaning that all evolutions that start in the same state must also end in the same state: evolutions can merge and cannot diverge.
+Therefore the evolution count, and therefore the entropy, can never decrease: $S(x(t)) \leq S(x(t+\Delta))$.
+A process is reversible if knowing $x(t + \Delta t)$ means we can reconstruct $x(t)$, meaning that all evolutions that end in the same state must also start in the same state: evolutions can diverge and cannot merge.
+If a process is deterministic end reversible, then, the evolution count, and therefore the entropy, must be conserved $S(x(t)) = S(x(t+\Delta))$.
+
+The last insight is that states are description of the system and only of the system.
+This happens only when the system is independent: the evolution of the system tells us nothing about the evolution of other systems.
+Independence means both factorization of the measure (and linearity of entropy) and it's maximization: the lack of correlations means all evolutions are possible.
+States can then be characterized by entropy in a way that is process independent, and lead to the geometrical structures of thermodynamics, classical, quantum and statistical mechanics.
 
@@
 The process entropy, then, can be expressed as $\int \rho \ln \rho$, and the Shannon entropy is recovered.
-
-TOADD: as the Boltzmann entropy fails to take into account interactions and correlations between molecules, the count of the microstates and the Shannon/Gibbs entropy fail to take into account system/environment correlations.
-Since these correlations may change in time, the same description of the same system will corresponds to different entropy in time.
 
@@
 We will not be discussing the Boltzmann entropy specifically, as this can be understood as a special case of the Gibbs entropy where all particles are identically and independently distributed.
+When correlations between particle/molecules are present, the Boltzmann entropy fails to correspond to the thermodynamic entropy.
 
@@
 we cannot isolate from gravitational interactions).
-Therefore entropy should be epistemological but objective: it quantifies the level of information (epistemological) that is accessible through a particular process allows (objective).
+Therefore entropy should be epistemological but objective: it quantifies the level of granularity/information (epistemological) that is accessible through a particular process allows (objective).
 
@@
 \textbf{Equilibria and relativity}.
-Another problem related to time dependence is that, in general theromdynamic equilibria are not relativistic object.
+Another problem related to time dependence is that the notion of theromdynamic equilibria is not, in general, relativistic.
 The notion of ``equilibria as nothing is changing'' is not invariant under boost.
@@
 The entropy of a single microstate is minus infinity, and not zero.
-A single microstate occupies zero entropy and the logarithm of zero is minus infinity.
+A single microstate occupies zero area and the logarithm of zero is minus infinity.
 In terms of Gibbs/Shannon entropy, a single microstate would correspond to a delta function.
@@
 \textbf{State space as template}.
-We should think of the state space as a template.
-We have a \textbf{state domain} $\tdomain_{X}$ which contain all statement fragments like \statement{the position of the ball is between 2 and 3 meters} together with a surjective map $\iota : \tdomain_{X} \times T \twoheadrightarrow \tdomain_t$ that adds \statement{at time t} for each possible time.
+We should think of the state space as a template we can use to instantiate groups of statements at each time.
+We have a \textbf{state domain} $\tdomain_{X}$ which contains all statement fragments like \statement{the position of the ball is between 2 and 3 meters} together with a surjective map $\iota : \tdomain_{X} \times T \twoheadrightarrow \tdomain_t$ that adds \statement{at time t} for each possible time.
 The map $\iota$ must be a surjection that preserves the logical structure (i.e.
-logical operations, verifiability, narrowness, compatibility, ...) so that the basic logical relationships determined by the system itself are are valid at all times (i.e.
+logical operations, verifiability, narrowness, compatibility, ...) so that the basic logical relationships determined by the system itself are valid at all times (i.e.
 \statement{at time t the position of the ball is between 2 and 3 meters} $\narrower$ \statement{at time t the position of the ball is between 0 and 300 meters} for every $t$).
@@
 As we said before, if the process is dissipative the set of possible states will become smaller.
-In that case, $\iota$ will map all the inaccessible states to $\contradiction$, as they will be impossible.
+In that case, $\iota$ will map all the inaccessible states to $\impossibility$, as they will be impossible.
 Therefore $\iota$ is not in general a bijection.
@@
 \end{itemize}
-There is a lot more work to be done to fully characterize this mathematical structure.
+There is a lot more work to be done to fully characterize this structure mathematically, but the current physical theory can serve as guidance.
+The point is that this structure has to serve as a general template that all physical theories must follow.
+We need a general definition of states and processes of which classical mechanics, quantum mechanics, thermodynamics and statistical mechanics are a specialization.
+The fact that the theories are, on the surface, so different is actually a good thing as it makes it easier to understand what elements are specific to the theories and what are more general.
 

### 2021-06-16 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: final pass from Andrea

@@
 \begin{abstract}
-\center Notes on classical and quantum logic
+It is generally believed that quantum mechanics entails a rejection of classical logic.
+The main arguments to support this claim come from the non-commutativity of quantum observables, that allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions.
+In this essay we show that such arguments contain serious flaws: on the one hand, it is possible to keep the classical distributivity law in the context of quantum theory once the actual meaning of quantum propositions is unambiguously defined and properly taken into account.
+On the other hand, we argue that the existence of quantum superpositions does not lead to the definition of a new rule for logical disjunctions.
+Furthermore, contrary to the received view in quantum logic, we explain that quantum mechanics can generate a distributive lattice of propositions.
+Finally, the remarkable formal similarities between the logical structures of classical and quantum mechanics are highlighted and discussed.
 \vspace{4mm}
 
-\noindent \emph{Keywords}: Quantum Mechanics; Quantum Logic; Classical Logic; 
+\noindent \emph{Keywords}: Quantum Mechanics; Quantum Logic; Classical Logic; Borel Algebra; Distributivity; Distributive Lattice
 \end{abstract}
@@
 As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus, a belief that is widespread still to this day (cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004} and \cite{deRonde:2016}).
-From the thirties, in fact, various modifications of Classical Logic (CL) were advanced in order to faithfully represent the physical content of quantum theory.
+From the thirties, indeed, various modifications of Classical Logic (CL) were advanced in order to faithfully represent the physical content of quantum theory.
 For instance, in 1931 the Polish philosopher Zygmunt Zawirski proposed to apply \L ukasiewicz's three-valued logic to QM, starting from considerations about the wave-particle duality and Heisenberg's uncertainty principle.
@@
 In their view, a typical example of invalid conjunction is one attributing a definite value for both position and momentum to a single quantum particle at a given time $t$.
-Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it is generally recognized that
-the founding text on Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that von Neumann in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 anticipated that from the algebraic structure of quantum theory it would have been possible to formulate a new propositional calculus.
+Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it is generally recognized that the founding text of Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that von Neumann in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 anticipated that from the algebraic structure of quantum theory it would have been possible to formulate a new propositional calculus.
 These ideas would have been fully expressed of his successive collaboration with Garrett Birkhoff.}
 
-A few years after the publication of this paper, the research in quantum logic had a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct formal structures to represent the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}).
+A few years after the publication of this paper, the research in quantum logic saw a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct logical structures to understand and describe the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}).
 In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography.
@@
 
-Recently one of us faced the question of whether quantum physics necessarily implies a revision of classical logic (cf.\ \cite{Oldofredi:2020}).
-Alternatively stated, it was asked if classical logic must be abandoned in the quantum realm, or there are quantum theories compatible with a classical propositional calculus.
-Analyzing Bohmian Mechanics, a well-known interpretation of QM, it has been showed that quantum physics does not necessarily involves a departure from classical logic.
-More precisely, it has been argued that in the context of this interpretation of QM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow classical rules---but also the distributivity law holds, contrary to the case of standard QM.
-This fact is a consequence of the clear ontology of the theory and its explanation of quantum measurements.
+Referring to this, one of us recently faced the question of whether quantum physics implies a revision of classical logic (cf.\ \cite{Oldofredi:2020}).
+Alternatively stated, it was asked if classical logic must be inevitably abandoned in the quantum realm, or if there are quantum theories compatible with classical propositional calculus.
+Analyzing Bohmian mechanics, a well-known interpretation of QM, it has been showed that quantum physics does not necessarily involves a departure from classical logic.
+More precisely, it has been argued that in this interpretation of QM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow classical rules---but also the distributivity law holds.
+This fact is a consequence of the particle ontology of the theory and its explanation of quantum measurements.
 
-In this paper we will be concerned with a similar issue, since we ask whether standard quantum mechanics necessarily involves a new logical structure radically different with respect to classical propositional calculus.
+In this paper we will be concerned with a similar issue, since we ask whether standard quantum mechanics itself necessarily involves a new logical structure radically different with respect to classical propositional calculus.
 In what follows we will answer this question in the negative.
-More precisely, here we have a two-fold aim: in the first place, we will analyze the main motivations and justifications for the introduction of quantum logic arguing that they do not entail the failure of CL in the quantum context.
-In the second place, we show that from the mathematical structures on which QM is built, it is possible to provide a distributive lattice to quantum propositions, contrary to the received view about QL.
-Such a distributive lattice of quantum statements, moreover, will allow us to take into account also assertions concerning expectation values, which are generally excluded in the context of the standard approach to quantum logic.
-Finally, we will argue that it is possible to demonstrate that classical logical propositions can form a non-distributive lattice.
+More precisely, this essay has a two-fold aim: in the first place, we analyze the main motivations and justifications for the introduction of quantum logic, claiming that such arguments do not really entail the failure of CL in the quantum context because they contain serious flaws.
+In the second place, we show that from the mathematical structures on which QM is built, it is possible to provide a distributive lattice of quantum propositions, contrary to the received view in QL.
+Such a distributive lattice of quantum statements, moreover, allows us to take into account assertions concerning expectation values, which are generally excluded in the standard approach to quantum logic.
+Finally, we argue that it is possible to demonstrate that classical logical propositions can form a non-distributive lattice.
 All these aspects, to our knowledge, are not sufficiently discussed in literature, thus, the present paper can contribute to shed new light on the relation between logic and quantum mechanics.
@@
 
-The essay is structured as follows: Section \ref{QL} provides a brief outline of the standard approach to QL (readers familiar with Birkhoff and von Neumann's proposal can skip this part), while in Section \ref{Motivations} we review the main arguments that have been given in literature to introduce a quantum logical calculus.
-Here we explain  that they do not entail the failure of classical logic in QM.
-In Section \ref{Math} we show not only that quantum logical propositions generate a distributive lattice, but also that important sentences of quantum theory---i.e.\ those concerning expectations values---are excluded by standard QL.\ In Section \ref{conc} we discuss the main philosophical implications of the preceding sections and conclude the paper.
+The essay is structured as follows: in Section \ref{QL} we review the main arguments that have been advanced in literature to introduce a quantum logical calculus, here we also provide a brief outline of the standard approach to QL (readers familiar with Birkhoff and von Neumann's proposal can skip this part).
+Section \ref{Motivations} explains why the main examples used to show the failure of classical logic in QM contain important flaws and ambiguities; moreover, we argue that the classical distributive law can be maintained in quantum mechanics, once such ambiguities are removed.
+In Section \ref{Math}, contrary to the received view in QL, we show not only that quantum logical propositions can generate a distributive lattice, but also that important sentences of quantum theory---i.e.\ those concerning expectations values---are excluded by standard quantum logic.\ In Section \ref{conc} we discuss the main philosophical implications of the preceding sections and conclude the paper.
 
@@
 
-Since part of the arguments in favor of QL stem from a claimed unsuitability of classical logic in the context of quantum mechanics, let us say a few words about classical logic, classical mechanics and their purported relationship.
+Given that a significant part of the arguments in favor of QL stem from a claimed unsuitability of classical logic in the context of quantum mechanics, let us say a few words about classical logic, classical mechanics and their purported relationship.
 It is a well known result that classical propositional logic is equivalent to a Boolean algebra where propositions (atomic and complex) have truth values ``true'' and ``false'' (or $\top$ and $\bot$), and the main operations among them are conjunction ``$\wedge$'', disjunction ``$\vee$'' and negation ``$\neg$'', the only unary operation.
@@
 It is also well established that the algebraic structure underlying classical mechanics is commutative, distributive and associative.
-The standard view in QL is that classical logic is applicable to classical mechanics because of the commutative algebraic structure.
+The standard view in QL is that classical logic is applicable to classical mechanics because of its commutative algebraic structure.
 
@@
 	\item the non-commutativity of quantum observables,
-	\item the peculiar behavior of the quantum disjunction---the second major deviation from classical logic---which may be true even though neither of the members is true in virtue of the linearity of the Schr\"odinger Equation (SE).
+	\item the peculiar behavior of the quantum disjunction---another major deviation from classical logic---which may be true even though neither of the members is true in virtue of the linearity of the Schr\"odinger Equation (SE).
 \end{itemize}
-\noindent Let us then have a closer look at these motivations for the failure of the distributivity law, and more generally of CL.
+\noindent Let us then have a closer look at these motivations for the failure of the distributivity law, and more generally of CL, in the quantum domain.
 
@@
 Furthermore, since spin operators along different axes are incompatible and do not commute---i.e.\ $[S_x, S_y]\neq 0, [S_x, S_z]\neq 0, [S_y, S_z]\neq0$---they represent incompatible observables that cannot be simultaneously measured.
-Let us then consider the following scenario: an electron is prepared with spin along the $x$ direction and subsequently the spin along the $y$ axis is measured.
-Let the propositions $q$, $p$ and $r$ be defined as follows:
+Let us then consider the following scenario: an electron is prepared with spin up along the $x$ direction and subsequently the spin along the $y$ axis is measured.
+Let the propositions $p$, $q$ and $r$ be defined as follows:
 \begin{description}
@@
 \end{description}
-As we said, there are only two states in which a particle can be found along an axis.
+As we said, there are only two states in which a particle can be found along a given axis.
 Therefore the proposition $q \vee r$ ``must be true'' (\cite{Giuntini:2002}, pp.\ 133-134).
@@
 \end{align}
-\noindent Since we assume that $p$ is true and we know that the disjunction $(q\vee r)$ is also true, whereas $(p\wedge q)$ and $(p\wedge r)$ are both false in virtue of the incompatibility between the spin operators among different axes.
+\noindent Since we assume that $p$ is true and we know that the disjunction $(q\vee r)$ is also true, we deduce that $p\wedge(q\vee r)=\top$; however, $(p\wedge q)$ and $(p\wedge r)$ are both false in virtue of the incompatibility between the spin operators among different axes.
 This fact implies that the right hand side of the distributive law asserts a false statement, i.e.\ $(p\wedge q)\vee(p\wedge r)=\bot$.
@@
 Their analysis begins by defining quantum logical propositions as \emph{experimental propositions}, i.e.\ statements affirming that a certain observable $A$ (or a set of observables) measured on a quantum system $s$ has a given value $a_i$ (or a sequence of values).
-More precisely, the authors stress that both in classical and quantum mechanics observations of physical systems are given by readings of experimental outcomes ($x_1, \dots, x_n$) of \emph{compatible} measurements ($\mu_1, \dots, \mu_n$).
+More precisely, the authors stress that both in classical and quantum mechanics observations of physical systems are given by the readings of the experimental outcomes ($x_1, \dots, x_n$) of \emph{compatible} measurements ($\mu_1, \dots, \mu_n$).
 The values ($x_1, \dots, x_n$) are elements of what the authors called the ($x_1, \dots, x_n$)-space, i.e.\ the ``observation-space'' of the system in question, whose elements are all the possible combinations of results of the compatible measurements ($\mu_1, \dots, \mu_n$).
@@
 A point in phase space corresponds to the specification of the position and momentum variables of a certain classical system, whereas in QM the points of $\mathcal{H}$ correspond to wave functions.
-The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcome in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.
-[TODO: check quote as it is grammatically incorrect]} as underlined by Svozi:
+The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcome in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.} as underlined by Svozi:
 \begin{quote}
@@
 
-In this section we will review the examples given in the previous one and see that the failure of distributivity stems from the ambiguity of the propositions and, once that is solved, standard classical rules can be applied.
-More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members is, and (ii) that \eqref{DL} can be retained in QM, since the arguments seen above---usually meant to show its failure---contain serious flaws.
+In this section we will review the examples given in the previous one and explain that the failure of distributivity stems from the ambiguity of the quantum logical propositions and, once that is clarified, standard classical rules can be applied also in the quantum context.
+More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members is, and (ii) that \eqref{DL} can be retained in QM, since the arguments seen above---usually meant to show its failure---are flawed.
 
@@
 As noted in the previous section, in classical mechanics one generally considers physical systems as objects having well-defined values for inherent, dynamical and relational properties which do not depend on any observation---meaning that measuring a magnitude on a classical particle will reveal a pre-existing value of the observed quantity.
-Thus, statements about the ontic state of classical systems before or after a measurement can be conflated without any dangerous metaphysical consequence.
+Thus, statements about the ontic state of classical systems before or after a measurement can be conflated without any dangerous metaphysical consequence.\footnote{Alternatively stated, in classical mechanics it is usually claimed that the sentence ``The value of the quantity $y$ for the system $s$ is $x$'' can be considered equivalent to the statement ``After a measurement $M$ of the quantity $y$ on the system $s$, the value $x$ is obtained''.
+This equivalence is due to the fact that measurements in the classical context do not alter the observed system, or in less idealized scenarios, that the disturbance caused by the interaction between observed system and measuring device is negligible.
+A more elaborate treatment of measurements in classical mechanics goes beyond the scope of the present paper.}
 
-However, in the context of QM this move is not allowed.
-The effect of the measurement on the state of the system, however one may choose to interpret them, cannot be disregarded.
-A statement about a system before a measurement, about its preparation\footnote{In operational quantum theory the preparations are a set of instructions that a certain agent has to follow in order to prepare a quantum system in a certain state.}, is different from a statement about the same system after its measurement, about the measurement outcome.
-In fact, it is only through repeated measurements that we can infer anything about the state of the system before the measurement.
-Thus, propositions like ``the electron has $y$-spin up'', ``the velocity of the particle is $v$'', ``the position of the particle is $x$'' are ambiguous as they do not spell out whether they are about preparation or measurement outcome.
-Once that is taken into account, the purported failure of classical logic disappears.
+However, in the context of QM this move is not allowed: the effect of a measurement on the state of a quantum system, however one may choose to interpret it, cannot be disregarded.
+A statement about a system before the performance of a certain measurement---i.e.\ about its preparation\footnote{In operational quantum theory the preparations are a set of instructions that an agent has to follow in order to prepare a quantum system in a certain state.
+We are aware that preparations can be considered special kinds of measurements.
+For the sake of the argument, however, in what follows we take preparations to be the instructions to arrange a physical system in a given state \emph{before} the measurement of a certain quantity is actually performed.}---is different from a statement regarding the same object when the measurement has been actually carried out, i.e.\ about the obtained measurement outcome.
+Indeed, it is only through repeated observations that we can infer anything about the state of the system before the measurement.
+Thus, propositions like ``the electron has $y$-spin up'', ``the velocity of the particle is $v$'', ``the position of the particle is $x$'' are inherently ambiguous as they do not spell out whether they are concerned with preparations or measurement outcomes.
+Once such vagueness is taken into account and clarified, the purported failure of classical logic disappears.
 
-Let us then redefine the propositions as follows:
+In order to support this claim, let us redefine the propositions encountered in the previous example as follows:
 \begin{description}
@@
 \end{description}
-\noindent Given the setup of the example, after the measurement either $q_o$ or $r_o$ must be true, which means $q_o \vee r_o = \top$.
-This reflects the simple fact that after a measurement of spin along the $y$ axis the state must be one of the two, which is true only after the measurement.
-The key is that $q_i$ and $r_i$ are not the same statements as $q_o$ and $r_o$, therefore it is not necessarily true that $q_o \vee r_o = q_i \vee r_i$.
-This unwarranted generalization is what causes the problem.
+\noindent Given the setup of the physical situation at hand, after the measurement of $S_y$ either $q_o$ or $r_o$ must be true, which means $q_o \vee r_o = \top$.
+This reflects the simple fact that \emph{after} a measurement of spin along the $y$ axis the state of the quantum particle must be one of the two indicated in either $q_o$ or $r_o$.
+Nonetheless, it is worth noting that $q_i$ and $r_i$ are \emph{not} the same statements as $q_o$ and $r_o$---i.e.\ $q_i$ and $r_i$ are not equivalent to $q_o$ and $r_o$---therefore, it is not necessarily true that $q_o \vee r_o = q_i \vee r_i$.
+This unwarranted equivalence is what causes the problem, to make evident why it is so, let us add a third step to our example.
 
+After the measurement of $S_y$, in the case we obtained $y$-spin down we will rotate the direction of spin by 90 degrees using the precession generated by a magnetic field, so that $x$-spin will be up, as we can see from Figure 1 below.
+However, in the other case, when $y$-spin up is obtained, we leave the direction of spin unchanged.
+
+\begin{figure}
 \tikzstyle{vecArrow} = [thick, decoration={markings,mark=at position
@@
 \end{center}
+\caption{\small{Graphic representation of the discussed example.
+Step 1 describes the preparation of the electron in the $x$-spin up state; Step 2 represents the measurement of $S_y$ at a later time; finally in Step 3 we rotate the direction of spin by 90 degrees if at Step 2 the particle is found in the $y$-spin down state.
+We leave the direction of spin unchanged otherwise.}}
+\end{figure}
 
-To make evident why the generalization is problematic, let us add a third step.
-After the measurement of $S_y$, in the case we obtained $y$-spin down we will rotate the direction of spin by 90 degrees using the precession generated by a magnetic field, so that $x$-spin will be up.
-However, in the other case, when $y$-spin up is obtained, we leave the direction of spin unchanged.
 Consider now the statements:
@@
 	\item $q_f$: ``after the third step, the electron $y$-spin is up'',
-	\item $r_f$: ``after the third step, the electron $y$-spin is down''.
+	\item $r_f$: ``after the third step, the electron $y$-spin is down'',
 \end{description}
-\noindent that assert the same properties as before but at the later time.
-By construction, the electron will either have $x$-spin up or $y$-spin up, either $p_f$ or $q_f$ must be true and therefore $p_f \vee q_f = \top$.\footnote{In fact, note that $q_f$ is true if and only if $q_o$ is true, and, similarly, $p_f$ is true if and only if $r_o$ is true.} But again, this relationship is valid only after the third step, which means $p_o \vee q_o$ is not necessarily true because these statements are evaluated at different times.
-But this situation is formally identical to the previous one: before we had that $q_o \vee r_o \neq q_i \vee r_i$, now we have $p_f \vee q_f \neq p_o \vee q_o$.
+\noindent that concern the state of the electron at the later time, after the possible rotation of spin.
+By construction, the particle will either have $x$-spin up or $y$-spin up, either $p_f$ or $q_f$ must be true and therefore $p_f \vee q_f = \top$.\footnote{In fact, note that $q_f$ is true if and only if $q_o$ is true, and, similarly, $p_f$ is true if and only if $r_o$ is true.} But again, this relationship is valid only after the third step, which means that $p_o \vee q_o$ is not necessarily true because these latter statements are evaluated at different (earlier) times.
+
+This situation, however, is formally identical to the previous one: before we had that $q_o \vee r_o \neq q_i \vee r_i$, now we have $p_f \vee q_f \neq p_o \vee q_o$.
 It would be improper to say that an electron has always either $x$-spin up or $y$-spin up just because it is so after the third step.
-In the same way, it would be improper to say that the electron as always either $y$-spin up or down just because it is so after the measurement of spin along the $y$ axis.
+Similarly, it would be improper as well to claim that the electron has always either $y$-spin up or down just because it is so after the measurement of spin along the $y$ axis.
 Neither the disjunction, nor its atomic components tell us anything about an inherent property associated with the observable $S_y$ possessed by a quantum particle independently of any measurement of $S_y$.
@@
 On the other hand, we can calculate the truth value of $q_i \vee r_i$ directly and show that it is false.
-The difficulty here is that before the measurement only $x$-spin is defined, so how are we going to compute statements about the value of $y$-spin?
-The solution is that while the value of an observable is not always defined, its expectation value always is.
+The difficulty here is that before the measurement of $S_y$ only $x$-spin is defined, so how are we going to compute statements about the value of $y$-spin?
+The solution is simple: while the value of an observable is not always defined, its expectation value always is.
 The idea is then to translate $q_i$ and $r_i$ into propositions about expectation values, which we already know how to handle, and apply the standard rules.
@@
 \begin{description}
-	\item $q'_i$: ``before the measurement, the expectation value for $y$-spin is $\frac{1}{2} \hbar$''
+	\item $q'_i$: ``before the measurement, the expectation value for $y$-spin is $\frac{1}{2} \hbar$'',
 	\item $r'_i$: ``before the measurement, the expectation value for $y$-spin is $-\frac{1}{2} \hbar$''.
@@
 Note that $q'_i \vee r'_i$  is the proposition ``before the measurement, the expectation value for $y$-spin is $\pm \frac{1}{2} \hbar$''.
-Now, before the measurement we know that the electron $x$-spin is up.
+Now, before the measurement of $S_y$ we know that the electron $x$-spin is up.
 While the $y$-spin value is not fully defined---the state is not an eigenstate of $S_y$---the expectation value is fully defined and it is zero.
@@
 
-While a full discussion goes beyond the scope of this work, we should mention that it is rather odd to claim that one of the reasons QM fails to follow classical logic is the linearity of the SE given that linear systems abound in classical mechanics.
+While a full discussion goes beyond the scope of this work, we should mention that it is rather odd to claim that one of the reasons QM fails to follow classical logic is the linearity of the SE given that linear systems abound in classical mechanics as well.
 All wave equations are linear which include both the SE and classical electromagnetism.
 The Liouville equation, the classical analogue to the SE that governs the evolution of probability distributions under classical Hamiltonian mechanics, is also linear.
-One can create superpositions of classical elecromagnetic fields or classical probability distributions in the same way one does of wave functions, and this does not seem to imply an invalidation of classical logic.
+One can create superpositions of classical electromagnetic fields or classical probability distributions in the same way one does of wave functions, and this does not seem to imply an invalidation of classical logic.
 
@@
 
-Let us now turn our attention to distributivity and try to rewrite \eqref{contradiction} by using the refined propositions that include the temporal dependence.
+Let us now turn our attention to distributivity and try to rewrite \eqref{contradiction} by using the refined propositions that explicitly include the temporal dependence.
 The proposition $p$ always referred to preparation, therefore it maps to $p_i$ in all cases.
-The disjunction was argued to be true on the basis of what happened after the measurement, therefore on the left side of the relationship we should use $q_o$ and $r_o$.
-On the right side, it was argued that $p \wedge q$ was false because of the incompatibility between the spin operators among different axis.
+The disjunction was argued to be true on the basis of what happened \emph{after} the measurement of $S_y$, therefore on the left side of the relationship we should use $q_o$ and $r_o$.
+On the right side, instead, it was argued that $p \wedge q$ was false because of the incompatibility between the spin operators among different axis.
 This means we are considering statements at equal time with $p_i$, and therefore we should use $q_i$ and $r_i$.
@@
 
-To sum up, \eqref{DL} fails because we considered propositions evaluated at different times, before and after the measurement.
+To sum up, \eqref{DL} fails in quantum domain because the we considered propositions evaluated at different times, before and after the measurement, in the two sides of the biconditional.
 If we remove the ambiguity and make the temporal dependence explicit, the distributivity law is recovered.
@@
 We can, in fact, construct a scenario in classical mechanics where propositions have the same before/after logical relationships.
-Suppose we have a ball sitting at position $p$ over a hatch (see Figure 1).
+Suppose we have a ball sitting at position $p$ over a hatch (see Figure 2).
 Once the hatch opens, the ball drops, bounces around, until it rests at either position $q$ or $r$ with equal chance.
@@
 \begin{description}
-    \item $p_i$: ``Before the hatch is opened, the position of the ball is $p$''
-    \item $q_i$: ``Before the hatch is opened, the position of the ball is $q$''
+    \item $p_i$: ``Before the hatch is opened, the position of the ball is $p$'',
+    \item $q_i$: ``Before the hatch is opened, the position of the ball is $q$'',
     \item $r_i$: ``Before the hatch is opened, the position of the ball is $r$''.
 \end{description}
-\noindent It is straightforward to see that these are three incompatible assertions: since the ball can have only one possible determinate position in space before the hatch is open, it cannot be localized in two different locations, $p$ and $q$ or $p$ and $r$.
+\noindent It is straightforward to see that these are three incompatible assertions: since the ball can have only one possible determinate position in space before the hatch is opened, it cannot be localized in two different locations, $p$ and $q$ or $p$ and $r$.
 Therefore the statements obey \eqref{solved_distrib_1}.
@@
 \begin{description}
-	\item $p_f$: ``After the hatch is opened, the position of the ball is $p$''
-	\item $q_f$: ``After the hatch is opened, the position of the ball is $q$''
-	\item $r_f$: ``After the hatch is opened, the position of the ball is $r$''.
+	\item $p_o$: ``After the hatch is opened, the position of the ball is $p$'',
+	\item $q_o$: ``After the hatch is opened, the position of the ball is $q$'',
+	\item $r_o$: ``After the hatch is opened, the position of the ball is $r$''.
 \end{description}
-Since the ball must be in one of the two lower positions after it dropped, the disjunction $(q_f \vee r_f) \equiv \top$.
+Since the ball must be in one of the two lower positions after it dropped, the disjunction $(q_o \vee r_o) \equiv \top$.
 Therefore the statements obey \eqref{solved_distrib_2}.
@@
 
-We can, however, evaluate $q$ and $r$ at different times, as it is done in the quantum case.
+We can, however, evaluate $q$ and $r$ at a different time w.r.t.\ $p$, as it is done in the quantum scenario considered in Section \ref{QL}.
 In that case, we obtain the same relationship as in \eqref{solved_contradiction}.
@@
 To this regard, it should be stressed that probabilistic theories are important even in classical mechanics.
-Due to the nature of experimental verification, also in the context of classical mechanics there are always going to be uncertainties in the preparation and measurements, and repeated measurements will directly lead to statistical constructs.
-However, we want to emphasize that this paper aims primarily at showing that QM can be given a distributive lattice of propositions.
-Here we want to enter neither into debates concerning the ontology of quantum mechanics, nor into discussions concerning the similarities between the commitments of QM and those of classical mechanics.
+Due to the nature of experimental verification, also in the context of classical mechanics there are always going to be uncertainties in the preparation and measurements, and repeated observations will directly lead to statistical constructs.
+However, we want to emphasize that this paper aims primarily at showing that QM can generate a distributive lattice of propositions.
+Here we want to enter neither into debates concerning the ontology of quantum mechanics, nor into discussions about the similarities between the commitments of QM and those of classical mechanics.
 In this essay we simply point out a formal similarity between the above mentioned mathematical structures that has implications for the logic of both theories.
@@
 We will claim that the non-distributive lattice is a strict subset of the distributive one; this result entails that everything that can be described in quantum logic can also be described in classical logic.
-Moreover, the distributive lattice will include propositions, like sentences concerning the expectation value of observables that are of high scientific interest, since in an experimental context by ``measurement'' one typically refers to the statistical collection of many ``single takes''.
+Moreover, the distributive lattice will include propositions concerning the expectation value of observables that are of high scientific interest, since in an experimental context by ``measurement'' one typically refers to the statistical collection of many ``single takes''.
 On the contrary, it is well-known that standard quantum logic deals only with single-take measurements, remaining silent w.r.t.\ propositions involving statistical statements.
@@
 Clearly, $\sigma$-algebras are Boolean algebras because they are exactly an algebra of sets with standard set operations.
-As we said before, the Borel algebra will contain the lattice of subspaces plus all the set operations between them, including the union.
+As we said before, the Borel algebra will contain the lattice of subspaces as well as all the set operations between them, including the union.
 In a way, the lattice of subspaces is not Boolean because it does not contain enough sets, which are instead contained in the Borel algebra.
@@
 	Let $\mathcal{H}$ be a Hilbert space and $\langle \, , \, \rangle$ be its inner product.
-Let $X$ be a self-adjoint bounded linear operator.
-Let $F_X : \mathcal{H} \to \mathbb{R}$ be the map defined by $F_X(\psi) = \langle \psi , X \psi \rangle$.
+Let $A$ be a self-adjoint bounded linear operator.
+Let $F_A : \mathcal{H} \to \mathbb{R}$ be the map defined by $F_A(\psi) = \langle \psi , A \psi \rangle$.
 Let $U \subseteq \mathbb{R}$ be a Borel set.
-Then $F_X^{-1}(U)$ is a Borel set.
+Then $F_A^{-1}(U)$ is a Borel set.
 \end{prop}
@@
 Proof.
-Let $X$ be a self-adjoint bounded linear operator.
+Let $A$ be a self-adjoint bounded linear operator.
 Recall that a linear operator is continuous if and only if it is bounded.
-Therefore $X$ is continuous.
-Note that $F_X$ is the composition of $X$ with the inner product.
+Therefore $A$ is continuous.
+Note that $F_A$ is the composition of $A$ with the inner product.
 Recall that the inner product is a continuous function.
-Since $F_X$ is the composition of continuous functions it is also continuous.
-Recall that all continuous functions are also Borel measurable, therefore $F_X$ is Borel measurable.
+Since $F_A$ is the composition of continuous functions it is also continuous.
+Recall that all continuous functions are also Borel measurable, therefore $F_A$ is Borel measurable.
 Let $U \subseteq \mathbb{R}$ be a Borel set.
-Since $F_X$ is Borel measurable, the reverse image $F_X^{-1}(U)$ is also a Borel set.
+Since $F_A$ is Borel measurable, the reverse image $F_A^{-1}(U)$ is also a Borel set.
 This concludes the proof.
 
-Since $\langle \psi , X \psi \rangle$ expresses the expected value for operator $X$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators.
+Since $\langle \psi , A \psi \rangle$ expresses the expected value for operator $A$, the proposition shows that the Borel algebra can express statements about the expectation value of observables,\footnote{Technically, position and momentum are not bounded operators.
 However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limited.
-Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity.
-Renormalization would lead to the set $\{ \psi \in H \, | \, \langle \psi , X \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $H \setminus \{ 0 \}$.}
+Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in H \, | \, \langle \psi , A \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity.
+Renormalization would lead to the set $\{ \psi \in H \, | \, \langle \psi , A \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $H \setminus \{ 0 \}$.}
 
@@
 That is, if the $x_0$ is the support of the wave-function.
-The former is a lot less stringent: the wave-function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$.
+The former is much less stringent: the wave-function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$.
 This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.\ a Gaussian wave packet can be centered on any value of position and momentum).
@@
 A state is identified by a two-valued distribution $[\rho_1, \rho_2]$, where $\rho_1$ is the probability to find the object in state 1 and $\rho_2$ is the probability to find the object in state 2.
-We note $[k,0]$ the subspace spanned by the first state only.
+We denote with $[k,0]$ the subspace spanned by the first state only.
 This corresponds to the statement ``the system $s$ is certainly in state $1$'' since the probability for state $2$ is zero.
@@
 It's a property of the lattice of subspaces of any vector space and, more generally, of the lattice of subgroups of any group.
-Indeed, as a further example, consider the group of boost transformations in relativity.
+Indeed, as a further example, consider the group of boost transformations in relativity theory.
 The boosts along one direction form a subgroup.
 If we take the subgroups of boost along $x$ and $y$ respectively, the smallest subgroup that contain both, the disjunction, is the subgroup of all boosts along any direction in the $x$-$y$ plane.
-As before, the disjunction is the span, so the resulting set is different from the union, and the union of all boosts along the x-axis and the y-axis is not a subgroup and therefore not contained in the lattice.
+As before, the disjunction is the span, hence the resulting set is different from the union, and the union of all boosts along the $x$-axis and the $y$-axis is not a subgroup and therefore not contained in the lattice.
 Thus, this has the same non-distributive structure as quantum logic.
 
-We want to be clear that we are not claiming that the lattice of subspaces of quantum logic is not of interest, or that it is of equal interest in both classical and quantum mechanics.
-Here we have pointed out that the argument that quantum mechanics \emph{requires} necessarily a radical departure from classical logic is untenable, given that both theories are built on similar structures.
-We conclude that such claims should be substantially weakened, in the light of the arguments that we proposed in this section.
+We want to stress once again that we are not claiming that the lattice of subspaces of quantum logic is not of interest, or that it is of equal interest in both classical and quantum mechanics.
+Here we have pointed out that the argument that quantum mechanics \emph{implies} a radical departure from classical logic is untenable, given that both theories are built on similar algebraic structures.
+We conclude that such claims, still believed to this day, should be substantially weakened, in the light of the arguments that we proposed in this section.
+
+\section{Concluding Remarks}
+\label{conc}
+
+In this essay we considered the question of whether QM entails a rejection of classical logic.
+After having reviewed the main arguments usually given in literature to support such a claim, and having introduced the standard approach to quantum logic, we showed not only that the classical distributivity law can be maintained in the context of quantum theory, but also that the alleged peculiar behavior of quantum disjunctions---generally ascribed to the existence of superpositions---is easily explained once the meaning of quantum propositions is unambiguously defined.
+
+More precisely, in this essay we argued that in order to understand the actual meaning of quantum propositions it is important to take into account whether they refer to preparations or measurements.
+Indeed, it is the temporal ambiguity present in the example discussed in Section \ref{QL} that led to the wrong conclusion according to which the distributive law is refuted by the physical content of standard quantum mechanics.
+Contrary to this claim, we argued that making explicit the temporal dependence of logical propositions is sufficient to recover \eqref{DL} in the quantum context.
+In addition, this result entails that the non-commutativity of quantum observables does not play any role in refuting or rejecting the distributivity law, meaning that such peculiar algebraic property of quantum operators do not change the rules of logic, dissolving one of the most important myths of QL.
+Similarly, we explained that the existence of superpositions does not entail per se that quantum disjunctions are true even in the case in which neither of its members is.
+
+Furthermore, in Section \ref{Math} it has been shown that taking into account the relevant and fundamental algebraic structures on which quantum mechanics is built, i.e.\ the Borel algebra, it is possible to demonstrate that quantum theory can generate a distributive lattice.
+On the other hand, we also argued that considering the space of probability distributions over phase space one can assign a non-distributive lattice to classical propositions.
+These facts indicate that the complexity of the Hilbert space characterizing quantum theory, as well as the non-commutativity of quantum observables do not entail any revision of the rules of classical logic in the quantum domain.
+Referring to this, going back to the very mathematical foundations of both classical and quantum theories, we saw that the Borel algebra, which is a Boolean algebra, characterizes the formal edifice on which both theories are built, and guarantees that both frameworks can be given a distributive lattice of logical propositions, showing as a consequence a remarkable similarity.
+The non-distributivity of quantum lattices originates from the simple algebraic fact that they do not employ standard set operations.
+However, this feature is a property of the lattice of subspaces of \emph{any} vector space, and of the lattice of subgroups of any group more generally.
+Hence, we can safely conclude that the mathematical structures on which quantum theory rest do not entail a rejection of classical logic.
 
@@
 If we take two intervals $U_x$ and $U_p$ in position and momentum, and form the statement ``the state is wholly in $U_x$ and $U_p$'', meaning that the vector components are zero outside of those regions, in the classical case we always find at least one $\rho(x,p)$ that satisfies the bounds, while in quantum mechanics, if the bounds violate the uncertainty principle, no suitable $\psi(x)$ can be found.
-The point is that the differences are in the content of the logical structure, on the particular relationships among the particular statements, not on the type of structure.
-
-\section{Discussion and Concluding Remarks}
-\label{conc}
+The point at stake in this paper---which is often not adequately recognized---is that the differences  are in the content of the logical structure of these theories, i.e.\ on the particular relationships that exist among the logical statements, and not on the type of structure.
 \vspace{5mm}

### 2021-06-16 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: minor edits from Gabriele

@@
 More precisely, this essay has a two-fold aim: in the first place, we analyze the main motivations and justifications for the introduction of quantum logic, claiming that such arguments do not really entail the failure of CL in the quantum context because they contain serious flaws.
-In the second place, we show that from the mathematical structures on which QM is built, it is possible to provide a distributive lattice of quantum propositions, contrary to the received view in QL.
-Such a distributive lattice of quantum statements, moreover, allows us to take into account assertions concerning expectation values, which are generally excluded in the standard approach to quantum logic.
-Finally, we argue that it is possible to demonstrate that classical logical propositions can form a non-distributive lattice.
+In the second place, we show that the mathematical structure on which QM is built already provides us with a distributive lattice of quantum propositions, contrary to the received view in QL.
+Such a distributive lattice of quantum statements, moreover, allows us to take into account assertions concerning expectation values, which are generally excluded in the standard approach to quantum logic but are nonetheless experimentally relevant.
+Finally, we argue that classical probabilistic theory are also equipped with an analogue non-distributive lattice.
 All these aspects, to our knowledge, are not sufficiently discussed in literature, thus, the present paper can contribute to shed new light on the relation between logic and quantum mechanics.
@@
 Section \ref{Motivations} explains why the main examples used to show the failure of classical logic in QM contain important flaws and ambiguities; moreover, we argue that the classical distributive law can be maintained in quantum mechanics, once such ambiguities are removed.
-In Section \ref{Math}, contrary to the received view in QL, we show not only that quantum logical propositions can generate a distributive lattice, but also that important sentences of quantum theory---i.e.\ those concerning expectations values---are excluded by standard quantum logic.\ In Section \ref{conc} we discuss the main philosophical implications of the preceding sections and conclude the paper.
+In Section \ref{Math}, contrary to the received view in QL, we show that quantum theory already posses a distributive lattice of propositions in the form of the Borel algebra, which includes not only the lattice of standard quantum logic, but also important sentences such as those concerning expectations values.
+We also show that classical probabilistic theory posses a formally analogue structure.
+In Section \ref{conc} we discuss the main philosophical implications of the preceding sections and conclude the paper.
 
@@
 
-This situation, however, is formally identical to the previous one: before we had that $q_o \vee r_o \neq q_i \vee r_i$, now we have $p_f \vee q_f \neq p_o \vee q_o$.
+This situation is formally identical to the previous one: before we had that $q_o \vee r_o \neq q_i \vee r_i$, now we have $p_f \vee q_f \neq p_o \vee q_o$.
 It would be improper to say that an electron has always either $x$-spin up or $y$-spin up just because it is so after the third step.
-Similarly, it would be improper as well to claim that the electron has always either $y$-spin up or down just because it is so after the measurement of spin along the $y$ axis.
+Similarly, it would be improper to claim that the electron has always either $y$-spin up or down just because it is so after the measurement of spin along the $y$ axis.
 Neither the disjunction, nor its atomic components tell us anything about an inherent property associated with the observable $S_y$ possessed by a quantum particle independently of any measurement of $S_y$.
@@
 
-To sum up, \eqref{DL} fails in quantum domain because the we considered propositions evaluated at different times, before and after the measurement, in the two sides of the biconditional.
+To sum up, \eqref{DL} fails in the quantum domain because we considered propositions evaluated at different times, before and after the measurement, in the two sides of the biconditional.
 If we remove the ambiguity and make the temporal dependence explicit, the distributivity law is recovered.
@@
 For a simpler example, $X$  could correspond to the set of the real numbers for a continuous quantity.} By definition, a $\sigma$-algebra is a collection of subsets of $X$ closed under countable union, countable intersection and complement with respect to $X$; the closure condition requires that  performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra.
-Interestingly, these sets represent statements concerning the objects described by $X$, i.e.\ quantum states and classical probability distributions.\footnote{In the quantum case this is exemplified also in Section \ref{QL}.} Then, in virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets.
+Interestingly, these sets can be understood to represent statements concerning the objects described by $X$, i.e.\ quantum states and classical probability distributions.\footnote{In the quantum case this is exemplified also in Section \ref{QL}.} Namely, if $U \subseteq X$, the statement ``$x$ is in $U$'' tells us that the state $x$ must one of those delimited by $U$.
+Then, in virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets.
 Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement.
@@
 
-While all of this may seem suspect to some reader, we would like to stress that these are very standard constructions.
+While all of this may seem suspect to some readers, we would like to stress that these are very standard constructions.
 Alternatively stated, all of what has been mentioned so far are the standard mathematical techniques used to study these spaces and can be found in any textbook (cf.\ \cite{Rudin:1991}, \cite{Vasudeva:2017}).

### 2021-06-19 Articles/2020-QuantumLogic/Source/quantumlogic.tex
Reviewed through end of Section 3.1

@@
 It is generally believed that quantum mechanics entails a rejection of classical logic.
-The main arguments to support this claim come from the non-commutativity of quantum observables, that allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions.
-In this essay we show that such arguments contain serious flaws: on the one hand, it is possible to keep the classical distributivity law in the context of quantum theory once the actual meaning of quantum propositions is unambiguously defined and properly taken into account.
-On the other hand, we argue that the existence of quantum superpositions does not lead to the definition of a new rule for logical disjunctions.
+The main arguments to support this claim come from the non-commutativity of quantum observables, which allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions.
+In this essay we show that such arguments contain serious flaws.
+In particular, it is possible to keep the classical distributivity law in the context of quantum theory once the actual meaning of quantum propositions is unambiguously defined and properly taken into account.
+We moreover argue that the existence of quantum superpositions does not lead to the definition of a new rule for logical disjunctions.
 Furthermore, contrary to the received view in quantum logic, we explain that quantum mechanics can generate a distributive lattice of propositions.
@@
 Furthermore, in 1933 Fritz Zwicky suggested that QM rejects the law of excluded-middle.\footnote{For historical details on quantum logic see \cite{Jammer:1974}.} However, it is generally recognized that the founding text of Quantum Logic (QL) is Birkhoff and von Neumann's essay ``\emph{The Logic of Quantum Mechanics}'' (\cite{vonNeumann:1936}), where the authors provided the standard account of the propositional calculus obeyed by quantum propositions.\footnote{It is worth noting that von Neumann in his treatise \emph{Mathematische Grundlagen der Quantenmechanik} published in 1932 anticipated that from the algebraic structure of quantum theory it would have been possible to formulate a new propositional calculus.
-These ideas would have been fully expressed of his successive collaboration with Garrett Birkhoff.}
+These ideas would be fully expressed in his successive collaboration with Garrett Birkhoff.}
 
@@
 In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography.
-Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now accepted among experts; for details see e.g.\ by \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
+Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
 Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory].
-This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).}, it is an important discipline of contemporary physics.
+This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).}, it remains an important discipline of contemporary physics.
 In this paper we will neither take a stance in the dispute concerning the ontological status of logic and its laws, nor we will consider recent developments of QL given that they are not strictly relevant to the point we are going to make.
-Indeed, the present essay is focused on the logical structures common to both quantum and Classical Mechanics (CM), following the spirit of the former discussion.\footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM.
-First order and higher order logics will not be discussed here as well as the logic of relativistic quantum theories.}
+Indeed, the present essay is focused on the logical structures common to both quantum and classical mechanics, following the spirit of the former discussion.\footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM.
+First-order and higher-order logics will not be discussed here, nor the logic of relativistic quantum theories.}
 
@@
 Alternatively stated, it was asked if classical logic must be inevitably abandoned in the quantum realm, or if there are quantum theories compatible with classical propositional calculus.
-Analyzing Bohmian mechanics, a well-known interpretation of QM, it has been showed that quantum physics does not necessarily involves a departure from classical logic.
+Analyzing Bohmian mechanics, a well-known interpretation of QM, it has been shown that quantum physics does not necessarily involve a departure from classical logic.
 More precisely, it has been argued that in this interpretation of QM not only the logical connectives retain their classical meanings---in particular the negation and the disjunction do follow classical rules---but also the distributivity law holds.
@@
 Such a distributive lattice of quantum statements, moreover, allows us to take into account assertions concerning expectation values, which are generally excluded in the standard approach to quantum logic but are nonetheless experimentally relevant.
-Finally, we argue that classical probabilistic theory are also equipped with an analogue non-distributive lattice.
-All these aspects, to our knowledge, are not sufficiently discussed in literature, thus, the present paper can contribute to shed new light on the relation between logic and quantum mechanics.
+Finally, we argue that classical probabilistic theory is also equipped with an analogue non-distributive lattice.
+All these aspects, to our knowledge, are not sufficiently discussed in literature; thus, the present paper can help to shed new light on the relation between logic and quantum mechanics.
 \vspace{2mm}
 
-The essay is structured as follows: in Section \ref{QL} we review the main arguments that have been advanced in literature to introduce a quantum logical calculus, here we also provide a brief outline of the standard approach to QL (readers familiar with Birkhoff and von Neumann's proposal can skip this part).
+The essay is structured as follows: in Section \ref{QL} we review the main arguments that have been advanced in the literature to introduce a quantum logical calculus; here we also provide a brief outline of the standard approach to QL (readers familiar with Birkhoff and von Neumann's proposal can skip this part).
 Section \ref{Motivations} explains why the main examples used to show the failure of classical logic in QM contain important flaws and ambiguities; moreover, we argue that the classical distributive law can be maintained in quantum mechanics, once such ambiguities are removed.
-In Section \ref{Math}, contrary to the received view in QL, we show that quantum theory already posses a distributive lattice of propositions in the form of the Borel algebra, which includes not only the lattice of standard quantum logic, but also important sentences such as those concerning expectations values.
-We also show that classical probabilistic theory posses a formally analogue structure.
+In Section \ref{Math}, contrary to the received view in QL, we show that quantum theory already possesses a distributive lattice of propositions in the form of the Borel algebra, which includes not only the lattice of standard quantum logic, but also important sentences such as those concerning expectations values.
+We also show that classical probabilistic theory possesses a formally analogue structure.
 In Section \ref{conc} we discuss the main philosophical implications of the preceding sections and conclude the paper.
@@
 79).
-To this regard, Beltrametti claims that 
+In this regard, Beltrametti claims that 
 \begin{quote}
@@
 
-Therefore, it is possible to formally characterize the state of a physical system via logical propositions which can assume the truth values ``true'' or ``false''---depending whether such statements describe true or false state of affairs concerning the system under consideration (cf.\ \cite{Jaeger:2009}, p.
+Therefore, it is possible to formally characterize the state of a physical system via logical propositions which can assume the truth values ``true'' or ``false''---depending whether such statements describe a true or false state of affairs concerning the system under consideration (cf.\ \cite{Jaeger:2009}, p.
 61)---and to perform logical operations among propositions via the connectives $(\wedge, \vee, \neg)$.
@@
 From a physical perspective this fact entails that if one performs a measurement of $A$ on a quantum system followed by a measurement of $B$, in general one will obtain different results inverting the order of these observations.
-Moreover, in virtue of the non-commutative algebra of quantum theory, to measure an operator $C=A+B$ is not typically equivalent to measure $A$ and $B$ independently and then add the respective results, since $A,B$ will be in general incompatible (cf.\ \cite{David:2015}, p.\ 77).
+Moreover, in virtue of the non-commutative algebra of quantum theory, measuring an operator $C=A+B$ is not typically equivalent to measuring $A$ and $B$ independently and then adding the respective results, since $A,B$ will be in general incompatible (cf.\ \cite{David:2015}, p.\ 77).
 This feature of QM constitutes the formal basis to prove the Heisenberg uncertainty relation, a theorem of quantum theory which reflects the operational inability to simultaneously prepare/measure the values of incompatible operators with arbitrary precision.
@@
 
-Let us consider a $1/2-$spin particle, such as an electron.
+Let us consider a spin-$1/2$ particle, such as an electron.
 From QM we know that there are only two possible spin states in which a particle can be found along each axis after a spin measurement, namely either in the up or in the down state.
@@
 A point in phase space corresponds to the specification of the position and momentum variables of a certain classical system, whereas in QM the points of $\mathcal{H}$ correspond to wave functions.
-The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcome in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.} as underlined by Svozi:
+The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcomes in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.} as underlined by Svozi:
 \begin{quote}
@@
 As a consequence, one is dealing with an operation that \emph{inverts} the two extreme probability-values, which naturally correspond to the truth-values \emph{truth} and \emph{falsity} (similarly to the classical truth-table of negation)'' (\cite{Giuntini:2002}, p.\ 132).
-\item Concerning the conjunction, Birkhoff and von Neumann notice that one can retain the very same set-theoretical interpretation of the classical conjunction in QL, since the intersection of two closed subspaces $P, Q\subset\mathcal{H}$ is still a closed subspace.
+\item Concerning the conjunction, Birkhoff and von Neumann note that one can retain the very same set-theoretical interpretation of the classical conjunction in QL, since the intersection of two closed subspaces $P, Q\subset\mathcal{H}$ is still a closed subspace.
 Thus, one maintains the usual meaning for the conjunction: the pure state $\psi$ verifies $P\cap Q$ iff $\psi$ verifies both $P$ and $Q$.
 Thus, quantum logic does not introduce a new logical operator for the conjunction.
-\item Contrary to the previous case, the logical operator for disjunction cannot be represented by the set-theoretic union, since the set-theoretic union of the two subspaces $P, Q$ will not be in general a subspace, thus, it is not an experimental proposition.
+\item Contrary to the previous case, the logical operator for disjunction cannot be represented by the set-theoretic union, since the set-theoretic union of the two subspaces $P, Q$ will not be in general a subspace; thus, it is not an experimental proposition.
 Therefore, in QL one introduces the quantum logical disjunction as the the closed span of the subspaces $P, Q$, which is an experimental proposition.
@@
 For the sake of the argument, however, in what follows we take preparations to be the instructions to arrange a physical system in a given state \emph{before} the measurement of a certain quantity is actually performed.}---is different from a statement regarding the same object when the measurement has been actually carried out, i.e.\ about the obtained measurement outcome.
-Indeed, it is only through repeated observations that we can infer anything about the state of the system before the measurement.
+Indeed, it is only through repeated observations on identically prepared systems that we can infer anything about the state of the system before the measurement.
 Thus, propositions like ``the electron has $y$-spin up'', ``the velocity of the particle is $v$'', ``the position of the particle is $x$'' are inherently ambiguous as they do not spell out whether they are concerned with preparations or measurement outcomes.
@@
 Nonetheless, it is worth noting that $q_i$ and $r_i$ are \emph{not} the same statements as $q_o$ and $r_o$---i.e.\ $q_i$ and $r_i$ are not equivalent to $q_o$ and $r_o$---therefore, it is not necessarily true that $q_o \vee r_o = q_i \vee r_i$.
-This unwarranted equivalence is what causes the problem, to make evident why it is so, let us add a third step to our example.
+This unwarranted equivalence is what causes the problem.
+To make evident why it is so, let us add a third step to our example.
 

### 2021-06-19 Articles/2020-QuantumLogic/Source/quantumlogic.tex
Reviewed through the end of Section 4.1

@@
 
-Before we saw that, if $p_i$ is true, the expectation for $y$-spin is zero, and therefore $q'_i \vee r'_i = q_i \vee r_i = \bot$. Therefore
+Before we saw that if $p_i$ is true, the expectation for $y$-spin is zero, and therefore $q'_i \vee r'_i = q_i \vee r_i = \bot$. Therefore
 \begin{align}
@@
 
+\begin{figure}
+
 \begin{center}
@@
 \end{center}
+\caption{\small{Graphic representation of the discussed example.
+Initially a ball is in position $p$ over a hatch.
+After the hatch opens, the ball drops and falls to either position $q$ or $r$ with equal chance.}}
+\end{figure}
 
@@
 In that case, we obtain the same relationship as in \eqref{solved_contradiction}.
-If we omit the indices, that is we are vague about the temporal evaluation, we get the same failure of distributivity as in \eqref{contradiction}.
-Naturally, one would not say that distributivity fails in the classical context, but rather that we have to evaluate propositions taking into account the same temporal instant in both sides of \eqref{DL}.
+If we omit the indices, that is, we are vague about the temporal evaluation, we get the same failure of distributivity as in \eqref{contradiction}.
+Naturally, one would not say that distributivity fails in the classical context, but rather that we have to evaluate propositions taking into account the same temporal instant on both sides of \eqref{DL}.
 
@@
 After having critically assessed the general motivations to introduce quantum logic as a consequence of the physical content of QM, in this section we present our main argument according to which it is possible to assign a distributive lattice to quantum propositions.
-Remarkably, in what follows we will see that the non-distributivity of the quantum logic lattice does neither originate from non-commuting observables nor from superpositions of quantum mechanics: it is a property of the lattice of subspaces of \emph{any} vector space, and of the lattice of subgroups of any group more in general (cf.\ \cite{Davey:2002}).
-To this regard, indeed, we also show that taking into account the space of classical probability distributions one can apply a non-distributive lattice to classical propositions.
+Remarkably, in what follows we will see that the non-distributivity of the quantum logic lattice originates neither from non-commuting observables nor from superpositions of quantum mechanics: it is a property of the lattice of subspaces of \emph{any} vector space, and of the lattice of subgroups of any group more in general (cf.\ \cite{Davey:2002}).
+In this regard, indeed, we also show that taking into account the space of classical probability distributions one can apply a non-distributive lattice to classical propositions.
 This fact is due to the similar structures of the state space of quantum mechanics and the space of classical probability distributions.
@@
 In what follows, then, it will be argued that both theories---quantum and classical---may be assigned \emph{two} lattices of propositions: a distributive one that follows the laws of classical logic, and a non-distributive one obeying the rules of quantum logic.
-Contrary to the widespread opinion, therefore, we will claim not only that quantum physics does not necessarily entails a revision of classical propositional calculus, but also that it is actually possible to show that quantum propositions can form a distributive lattice.
-Furthermore, in agreement with the example of the violation of the distributivity law in the classical context seen a few lines above, in the remainder of the section we will argue also that the propositions of classical mechanics may obey a non-classical type of logic.
+Contrary to the widespread opinion, therefore, we will claim not only that quantum physics does not necessarily entail a revision of classical propositional calculus, but also that it is actually possible to show that quantum propositions can form a distributive lattice.
+Furthermore, in agreement with the example of the violation of the distributivity law in the classical context seen above, in the remainder of the section we will argue also that the propositions of classical mechanics may obey a non-classical type of logic.
 
 Before we proceed, it is worth noting that the classical space we will be using is not phase-space itself, but the space of probability distributions over phase-space, which is a real Hilbert space and thus has a similar structure to its quantum counterpart.
-To this regard, it should be stressed that probabilistic theories are important even in classical mechanics.
+It should be stressed that probabilistic theories are important even in classical mechanics.
 Due to the nature of experimental verification, also in the context of classical mechanics there are always going to be uncertainties in the preparation and measurements, and repeated observations will directly lead to statistical constructs.
@@
 
-From a mathematical perspective, the distributive lattice will be provided in both classical and quantum case by a Borel algebra, which is the smallest $\sigma$-algebra on a set $X$ equipped with a topological structure that contains the topology; the non-distributive lattice will be provided by the set of closed subgroups ordered by inclusion.
+From a mathematical perspective, the distributive lattice will be provided in both the classical and the quantum case by a Borel algebra, which is the smallest $\sigma$-algebra on a set $X$ equipped with a topological structure that contains the topology; the non-distributive lattice will be provided by the set of closed subgroups ordered by inclusion.
 We will claim that the non-distributive lattice is a strict subset of the distributive one; this result entails that everything that can be described in quantum logic can also be described in classical logic.
@@
 For a simpler example, $X$  could correspond to the set of the real numbers for a continuous quantity.} By definition, a $\sigma$-algebra is a collection of subsets of $X$ closed under countable union, countable intersection and complement with respect to $X$; the closure condition requires that  performing those operations on the sets contained in a $\sigma$-algebra will return another set in the same $\sigma$-algebra.
-Interestingly, these sets can be understood to represent statements concerning the objects described by $X$, i.e.\ quantum states and classical probability distributions.\footnote{In the quantum case this is exemplified also in Section \ref{QL}.} Namely, if $U \subseteq X$, the statement ``$x$ is in $U$'' tells us that the state $x$ must one of those delimited by $U$.
-Then, in virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets.
-Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement.
+Interestingly, these sets can be understood to represent statements concerning the objects described by $X$, i.e.\ quantum states and classical probability distributions.\footnote{In the quantum case this is exemplified also in Section \ref{QL}.} Namely, if $U \subseteq X$, the statement ``$x$ is in $U$'' tells us that the state $x$ must be one of those delimited by $U$.
+Then, by virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets.
+Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement [TODO: For statements, should this be complement or negation?].
 Hence, the $\sigma$-algebra generates a Boolean algebra of statements.\footnote{Note that the interpretation used here is exactly the same used in probability theory.
@@
 
-In the second place, it is crucial to underline that every topological space---as for instance the complex Hilbert space of quantum mechanics and the real Hilbert space of classical probability distributions considered in this essay---comes already equipped with a $\sigma$-algebra called Borel algebra.\ A topology, in fact, is also a collection of sets---called \emph{open sets}---and the Borel algebra is defined to be the smallest $\sigma$-algebra that contains all the open sets.\footnote{Here the topology is important for us only as a mean to construct the Borel algebra.} Complete normed spaces, such as the mentioned Hilbert spaces, have a canonical way to construct a topology and, therefore, a $\sigma$-algebra.
+In the second place, it is crucial to underline that every topological space---as for instance the complex Hilbert space of quantum mechanics and the real Hilbert space of classical probability distributions considered in this essay---comes already equipped with a $\sigma$-algebra called a Borel algebra.\ A topology, in fact, is also a collection of sets---called \emph{open sets}---and the Borel algebra is defined to be the smallest $\sigma$-algebra that contains all the open sets.\footnote{Here the topology is important for us only as a means to construct the Borel algebra.} Complete normed spaces, such as the mentioned Hilbert spaces, have a canonical way to construct a topology and, therefore, a $\sigma$-algebra.
 Let us see how this procedure works in the easier case of a standard three-dimensional Euclidean space.
@@
 We can define a distance $\epsilon^2 = |\psi - \phi|^2=\langle \psi - \phi , \psi - \phi \rangle$ between two states, and the statement ``the state of the system is within $\epsilon$ of $\psi$'' corresponds to the set $U = \{\phi \; | \; |\psi - \phi|^2 < \epsilon^2\}$.
-As we have already stated, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.\footnote{The choice of the square root of the density over the density itself is merely a technical choice.} In this case we have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
+As we have already stated, the space of classical probability distributions is also a Hilbert space where each point corresponds to the square root of the density.\footnote{The choice of the square root of the density rather than the density itself is merely a technical choice.} In this case we have a distance $\epsilon^2 = |\sqrt{\rho_1} - \sqrt{\rho_2}|^2=\langle \sqrt{\rho_1} - \sqrt{\rho_2} , \sqrt{\rho_1} - \sqrt{\rho_2} \rangle$ and we can proceed in the same exact way.
 In all normed spaces, the topology, and therefore the $\sigma$-algebra, is constructed from sets of the form ``the object is within $\epsilon$ of $x$''.
@@
 
-Arrived at this point, one may ask: do these extra sets correspond to physically meaningful propositions? To answer this question, let us first demonstrate the following proposition.
+At this point, one may ask: do these extra sets correspond to physically meaningful propositions? To answer this question, let us first demonstrate the following proposition.
 
@@
 Borel sets also allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
-This is \emph{not} equivalent to the statement ``the particle is in the eigenstate $x_0$''.
+This is \emph{not} equivalent to the statement ``the particle is in the eigenstate $x_0$'' [TODO: Does it make sense to say that the eigenstate, rather than the eigenvalue, is $x_0$?
+Is different notation needed to make it a state?].
 This latter proposition will be true only if the whole function is at $x_0$.
-That is, if the $x_0$ is the support of the wave-function.
-The former is much less stringent: the wave-function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$.
+That is, if $x_0$ is the support of the wave function.
+The former is much less stringent: the wave function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$.
 This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.\ a Gaussian wave packet can be centered on any value of position and momentum).
@@
 Recall that the Ehrenfest theorem states that $m \frac{d}{dt}\langle x \rangle = \langle p \rangle$ and $ \frac{d}{dt}\langle p \rangle = - \langle V(x) \rangle$.
-If the potential for a quantum particle can be considered constant over the wave-function (i.e.
+If the potential for a quantum particle can be considered constant over the wave function (i.e.
 $\langle V(x) \rangle = V(\langle x \rangle)$ ), the motion of the center of mass will reduce to the case of a point-particle.

### 2021-06-21 Articles/2020-QuantumLogic/Source/quantumlogic.tex
Finished review

@@
 
-Now that we have argued that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
+Now that we have argued that quantum mechanics can be given an algebra that looks more like [TODO: Should a stronger phrase than "looks more like" be used?] the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
 Given that quantum mechanics is inherently a probabilistic theory, let us compare it to a classical probabilistic theory.
@@
 
-Note that during the discussion there was no mention as to whether $k$ and $j$ where real numbers, complex numbers or quaternions.
+Note that during the discussion there was no mention as to whether $k$ and $j$ were real numbers, complex numbers or quaternions.
 The argument is formally the same in all cases.
@@
 Contrary to this claim, we argued that making explicit the temporal dependence of logical propositions is sufficient to recover \eqref{DL} in the quantum context.
-In addition, this result entails that the non-commutativity of quantum observables does not play any role in refuting or rejecting the distributivity law, meaning that such peculiar algebraic property of quantum operators do not change the rules of logic, dissolving one of the most important myths of QL.
+In addition, this result entails that the non-commutativity of quantum observables does not play any role in refuting or rejecting the distributivity law, meaning that this peculiar algebraic property of quantum operators does not change the rules of logic, dissolving one of the most important myths of QL.
 Similarly, we explained that the existence of superpositions does not entail per se that quantum disjunctions are true even in the case in which neither of its members is.
@@
 Furthermore, in Section \ref{Math} it has been shown that taking into account the relevant and fundamental algebraic structures on which quantum mechanics is built, i.e.\ the Borel algebra, it is possible to demonstrate that quantum theory can generate a distributive lattice.
-On the other hand, we also argued that considering the space of probability distributions over phase space one can assign a non-distributive lattice to classical propositions.
-These facts indicate that the complexity of the Hilbert space characterizing quantum theory, as well as the non-commutativity of quantum observables do not entail any revision of the rules of classical logic in the quantum domain.
+Moreover, we argued that considering the space of probability distributions over phase space one can assign a non-distributive lattice to classical propositions.
+These facts indicate that neither the complexity of the Hilbert space characterizing quantum theory nor the non-commutativity of quantum observables entails any revision of the rules of classical logic in the quantum domain.
 Referring to this, going back to the very mathematical foundations of both classical and quantum theories, we saw that the Borel algebra, which is a Boolean algebra, characterizes the formal edifice on which both theories are built, and guarantees that both frameworks can be given a distributive lattice of logical propositions, showing as a consequence a remarkable similarity.

### 2021-06-21 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: addressing Christine's TODO

@@
 Then, by virtue of the equivalence between sets and statements, it will be more convenient for us to conceive $\sigma$-algebras as collections of propositions about $X$ instead of sets.
-Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and complement [TODO: For statements, should this be complement or negation?].
+Referring to this, since the $\sigma$-algebra is closed under countable union, countable intersection and complement with respect to $X$, the corresponding statements will be closed under countable conjunction, countable disjunction and negation.
 Hence, the $\sigma$-algebra generates a Boolean algebra of statements.\footnote{Note that the interpretation used here is exactly the same used in probability theory.
@@
 Borel sets also allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
-This is \emph{not} equivalent to the statement ``the particle is in the eigenstate $x_0$'' [TODO: Does it make sense to say that the eigenstate, rather than the eigenvalue, is $x_0$?
-Is different notation needed to make it a state?].
-This latter proposition will be true only if the whole function is at $x_0$.
+This is \emph{not} equivalent to the statement ``the particle is in the eigenstate of position corresponding to $x_0$'' This latter proposition will be true only if the whole function is at $x_0$.
 That is, if $x_0$ is the support of the wave function.
@@
 
-Now that we have argued that quantum mechanics can be given an algebra that looks more like [TODO: Should a stronger phrase than "looks more like" be used?] the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
+Now that we have argued that quantum mechanics can be given an algebra that looks more like the one of classical mechanics, let us show that we can (at least formally) give classical mechanics an algebra that looks like the one of quantum logic.
 Given that quantum mechanics is inherently a probabilistic theory, let us compare it to a classical probabilistic theory.

### 2021-07-02 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: adding example after Gordon's feedback

@@
 However, one can argue that measured position and momentum are indeed bounded as the acceptance of a detector is always limited.
-Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in H \, | \, \langle \psi , A \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity.
-Renormalization would lead to the set $\{ \psi \in H \, | \, \langle \psi , A \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $H \setminus \{ 0 \}$.}
+Note that all operators for which the spectral theorem applies can be written as the limit of a sequence of bounded operators.} such as ``the expected position $\langle x \rangle$ of the particle is between $x_0$ and $x_1$.'' This would correspond to the Borel set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle \in (x_0, x_1) \}$.\footnote{In most cases, one assumes states to be normalized, which is what we have done here for simplicity.
+Renormalization would lead to the set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $\mathcal{H} \setminus \{ 0 \}$.} Consider, for example, ``the expectation of $z$-spin is zero''.
+This, by Proposition 2, would correspond to a Borel set $V \subseteq \mathcal{H}$.
+This set is not the full set since the state with $z$-spin up is not in the set.
+However, $x$-spin up and $x$-spin down will be in the set, and their span is the full space.
+Therefore $V$ is not a closed subspace.
+Our proposition, despite being physically interesting, does not correspond to a closed subspace.
 
-Borel sets also allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
+The idea is that closed linear subspaces correspond to single-shot measurements.
+While these form an interesting subset of possible measurements, they are fairly limited in what they can tell us about the incoming system.
+For example, measuring $z$-spin up tells us very little about spin \textbf{before} the measurement: only that it couldn't be $z$-spin down.
+To make any inference on the incoming state we need to gather statistics.
+In fact, no cross section measurement, our main tool in exploring the standard model, is single-shot.
+Being that quantum mechanics is a statistical theory, most physically interesting statements are statistical in nature, and that is why we need to include the Borel sets in our lattice of statements.
+
+Another distinction is important.
+Borel sets allow us to express statements such as ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''.
 This is \emph{not} equivalent to the statement ``the particle is in the eigenstate of position corresponding to $x_0$'' This latter proposition will be true only if the whole function is at $x_0$.
 That is, if $x_0$ is the support of the wave function.
-The former is much less stringent: the wave function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long the average is $x_0$.
+The former is much less stringent: the wave function can even be distributed over an infinite range (e.g.\ a Gaussian wave packet) just as long as the average is $x_0$.
 This means that we can have statements like ``the expected position $\langle x \rangle$ of the particle is exactly $x_0$''$\wedge$ ``the expected momentum $\langle p \rangle$ of the particle is exactly $p_0$'' without having contradictions (e.g.\ a Gaussian wave packet can be centered on any value of position and momentum).

### 2021-07-12 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
ClassMechPhil: adding disclaimer paragraph based on feedback of rev 2

@@
 This shows that Lagrangian systems are a proper subset of Hamiltonian systems, elucidating the connection between these two formalisms.
-Using our framework, we furthermore argue that the invariant phase space density is physically fundamental.
+Moreover, we argue that the invariant phase space density is physically fundamental.
 Since Hamiltonian mechanics provides the proper setting for this density, we gain another reason for privileging it over Lagrangian mechanics.
@@
 
-Additionally, whereas North \parencites*[76]{North} \textbf{NOTE: before the citation was just \textbf{North} without the repetition} seeks to determine the `real' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure ontologically, but as a formal device for encoding physical information or a calculational tool.
+Additionally, whereas North \parencites*[76]{North} seeks to determine the `real' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure ontologically, but as a formal device for encoding physical information or a calculational tool.
 As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is `fundamental in the bottom-layer sense' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental in a classical world does not matter for the questions we seek to answer.
@@
 We thus demonstrate in detail our methodology of deriving  relevant mathematical structures from underlying physical conditions.
+
+Before starting, we want to make it clear that the following is not and cannot be a formal proof.
+The idea is that the assumptions put constraints on the physical entities and, in turn, on the mathematical objects that are used to represent them.
+These constraints are constitutive in nature, they provide requirements for the math, and therefore they cannot be derived from the math.
+Each assumption will introduce multiple constraints, which will be introduced gradually within each section.
+The goal is to show how such constraints follow logically from the assumptions, how they are encoded in the math and how they recover the well known structures.
 
@@
 In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
-If the value of the density at one point is larger than another, physically it means that there is more matter at this point.
+If the value of the density at one point is larger than another, then physically there is more matter at this point.
 Likewise, if the value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
-For brevity, we'll refer to this as allowing \textit{point-wise comparison}.
-Point-wise comparison requires a coordinate-invariant density.
+For brevity, we'll refer to this as the property of \textit{point-wise comparability}.
+Point-wise comparability requires a coordinate-invariant density.
 As the derivation in Section~\ref{infinitesimal} shows, to guarantee the existence of an invariant density, we must define state space over a symplectic manifold.
-In other words, to generically have a density which allows point-wise comparison, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
-
-Section~\ref{dependence} shows why point-wise comparison does not hold in general: the density is a coordinate-dependent quantity, and thus it is not generically coordinate-invariant.
-Within the Lagrangian formalism, point-wise comparison holds only when we have made a special choice of coordinates.
+In other words, to generically have a point-wise comparable density, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
+
+Section~\ref{dependence} shows why point-wise comparability does not hold in general: the density is a coordinate-dependent quantity, and thus it is not generically coordinate-invariant.
+Within the Lagrangian formalism, point-wise comparability holds only when we have made a special choice of coordinates.
 This gives a further reason to privilege Hamiltonian mechanics.
@@
 
-\subsection{Point-wise comparison}
+\subsection{Point-wise comparability}
 \label{dependence}
@@
 Since the Jacobian is a function of the point, the value of $|J|$ can change from point to point.
-Consequently, the density is not guaranteed to be directly comparable at different points: point-wise comparison need not hold in an arbitrary coordinate system.
+Consequently, the density is not guaranteed to be directly comparable at different points: point-wise comparability need not hold in an arbitrary coordinate system.
 If we have a uniform, constant density in $x$ and $y$, we will have an $r$-dependence in $r$ and $\theta$.
@@
 The result is that we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in (in the sense of space-time coordinates).
-Any and all (time-independent) canonical coordinate systems will support point-wise comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space $(q^i, p_i, t, E)$.
-See \textcites[]{Struckmeier}.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that allow for point-wise comparison of the density.\footnote{One might wonder if it is possible to construct a coordinate-independent density, perhaps using either (i) volume forms from differential geometry or (ii) the canonical volume form of Riemannian geometry.
+Any and all (time-independent) canonical coordinate systems support point-wise density comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space $(q^i, p_i, t, E)$.
+See \textcites[]{Struckmeier}.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that support point-wise comparability.\footnote{One might wonder if it is possible to construct a coordinate-independent density, perhaps using either (i) volume forms from differential geometry or (ii) the canonical volume form of Riemannian geometry.
 Both of these strategies fail, although space precludes a demonstration.}
@@
 Moving beyond classical mechanics, our framework has the advantage of providing an elegant relationship with both classical statistical mechanics and quantum mechanics.
-We note that the interpretation of the symplectic form is exactly the one used by the fundamental postulate of statistical mechanics, which uses phase space volumes (i.e.
-the volume form generated by the symplectic form) to count states.
-Additionally, by changing the requirement of infinitesimal reducibility to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.\footnote{We provide details of the quantum case elsewhere.} Hence, our framework perspicuously distinguishes quantum from classical systems.
-We take these considerations as reason enough for adopting our preferred notion of a classical system.
+First, the interpretation of the symplectic form is exactly the one used by the fundamental postulate of statistical mechanics, which uses phase space volumes---i.e.
+the volume form generated by the symplectic form---to count states.
+Second, by changing the requirement of infinitesimal reducibility to a corresponding notion of \textit{irreducibility}, we recover quantum mechanical systems.
+Hence, our framework perspicuously distinguishes quantum from classical systems.
+We take these considerations as reason enough for adopting our preferred notion of a classical system.\footnote{We provide details of the quantum and statistical mechanical cases elsewhere.}
 % % %, see \textcites{AoPPhy1}. [omit for review].
@@
 Finally, we have provided a geometrical interpretation of the action principle that makes the gauge-dependency of the Lagrangian manifest.
-Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but classical systems in general.
+Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but of classical systems in general.
 %The importance of the density shows that a quantity can be fundamental yet coordinate-dependent.

### 2021-07-26 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
Update classical-phil-article-BJPS.tex
Addressed most referee comments; some outstanding `to do's', especially in section 4.2.

@@
 Section~\ref{density} explains how the invariant phase space density plays a special role in the Hamiltonian framework.
-The density allows us to compare and track the amount of matter in different regions of phase space.
+The density allows us to compare and track the amount of matter in different possible states.
 In contrast, Lagrangian mechanics does not support these important properties of the density.
@@
 
-In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by `fundamental.' Of the various notions of fundamentality, we will mean something entirely non-mysterious: a formalism is more fundamental provided that it is logically more general.
-In this case, the Lagrangian framework is less general because it requires an additional condition to hold, compared to the Hamiltonian framework.
-Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.\footnote{One might worry that our criterion for fundamentality is subject to elementary counterexamples.
+In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by `fundamental.' Of various notions of fundamentality, we mean something entirely non-mysterious.
+Given two formalisms that capture a class of systems, if one is logically more general, then it is conceptually more fundamental.\footnote{In Section~\ref{density}, we embrace a related non-metaphysical notion of fundamentality that pertains to physical entities rather than theoretical frameworks.} Conceptual fundamentality is familiar from mathematics.
+A more fundamental framework encompasses a less fundamental one as a special case.
+It is in this sense that the generalized Stokes theorem is more fundamental than well-known special cases, such as the fundamental theorem of calculus and Green's theorem.
+Similarly, an axiomatization which includes additional axioms is less fundamental than one which drops these axioms while maintaining the same expressive power.
+Conceptual fundamentality is distinct from notions of fundamentality associated with differences in energy scale, such as how the Standard Model of particle physics is more fundamental than classical mechanics.
+We do not mean that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in this sense.
+
+To apply our criterion for fundamentality, we must show that Lagrangian mechanics is less general than Hamiltonian mechanics.
+We do this in Section~\ref{derivation} by recovering the mathematical formalisms of both Hamiltonian and Lagrangian mechanics from three physical conditions.
+Our aim here is to show how this argument works in principle.
+Section~\ref{fundamentality} demonstrates some philosophical advantages of our preferred notion of fundamentality.\footnote{One might worry that our criterion for fundamentality is subject to elementary counterexamples.
 For instance, can't we make any theory more general by adjoining arbitrary disjunctions?
@@
 Hamiltonian mechanics is more general than Lagrangian mechanics while still being about classical systems, where `aboutness' can be understood using Yablo's framework \parencites*[]{Yablo}.
-Having an account of \textit{aboutness} or \textit{subject matter} is a general philosophical problem that any account of fundamentality or essence must address.
-Hence, it is not a special problem facing our criterion.}
-
-Our argument has three main steps, summarized in Figure~\ref{diagram}.
-The left column lists three basic physical conditions: infinitesimal reducibility, determinism/reversibility, and motion equivalence.
-Each condition entails physically significant properties, described in the middle column.
-Each physical property corresponds to a formal/mathematical structure depicted in the right column, including phase space, Hamiltonian mechanics, and Lagrangian mechanics.
-In this way, our framework develops a correspondence between physical conditions and their associated mathematical formalisms.
-We thereby gain a straightforward interpretation of the physical significance of these formalisms.
-
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=\textwidth]{Diagram.png}
-\caption{Physical conditions and corresponding properties and formalisms}
-\label{diagram}
-\end{figure}
-
-First, a system is \textit{infinitesimally reducible} provided that the state of the whole system is equivalent to specifying the states of its parts (and vice versa).
+Having an account of \textit{aboutness} or \textit{subject matter} is a general philosophical problem that any account of fundamentality must address.
+It is not a special problem facing our criterion.}
+%We will show that the Lagrangian framework is less general because it requires an additional condition to hold, compared to the Hamiltonian framework.
+Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.
+
+%We will, however in Section 5 give a criterion for an entity to be physically fundamental
+%These bolster our claim that this is an appropriate sense of fundamentality for philosophers to focus on.
+
+Shorn of many details, our argument goes as follows.
+First, we show that the formalism of Hamiltonian mechanics applies to systems that are both i) infinitesimally reducible and ii) follow a deterministic and reversible evolution.
+A system is \textit{infinitesimally reducible} provided that the state of the whole system is equivalent to specifying the states of its parts (and vice versa).
+Colloquially, we can say that the whole state equals the sum of its parts, where we take these infinitesimal parts to be classical particles.
 Infinitesimal reducibility distinguishes classical from quantum systems, where the latter are irreducible (i.e.
 we cannot specify the whole state by specifying the states of its subsystems).
-In the widest sense, a classical system is any infinitesimally reducible system.
-Colloquially, we can say that the whole state equals the sum of its parts, where we take these infinitesimal parts to be classical particles.
-
-If a system is infinitesimally reducible, then we can define a distribution $\rho$ over particle states and a density $\rho_{\xi}$ over the corresponding state variables $\xi$.
-Since particle states are independent of a choice of units, this density must be invariant.
-We show in Section~\ref{infinitesimal} that defining this invariant density requires a symplectic manifold, which serves as the phase space of classical mechanics.
-The physical significance of this symplectic manifold is thereby its connection with the physical condition of infinitesimal reducibility.
-
-Secondly, a system has \textit{deterministic and reversible evolution} provided that the present state of the system uniquely determines both its future and past states.
-This condition entails that the invariant density $\rho_{\xi}$ is a conserved phase space density.
-We will show that a classical system which satisfies this condition is Hamiltonian: its phase space is a symplectic manifold that evolves according to Hamiltonian evolution  (i.e.
-according to a symplectomorphism).
+We show in Section~\ref{infinitesimal} that the phase space of infinitesimally reducible systems must have at least the structure of a symplectic manifold.
+Recovering Hamiltonian mechanics then simply requires identifying when trajectories on phase space obey Hamiltonian evolution  (i.e.
+evolve according to a symplectomorphism).
+Section~\ref{deterministic} shows that this requires the system to follow a \textit{deterministic and reversible evolution}, i.e.
+the present state of the system uniquely determines both its future and past states.
 We thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution.
-Section~\ref{deterministic} justifies this correspondence.
-
-Finally, some classical systems satisfy a further condition, which we call \textit{motion equivalence}.
+
+Next, we note that some classical systems satisfy a third, further condition, which we call \textit{motion equivalence}.
 Motion equivalence holds when the system's trajectory in physical space determines---and is determined by---its evolution in phase space.
@@
 As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
-
-We cannot always determine the evolution of a system in phase space from its trajectory in physical space, since some systems fail to satisfy motion equivalence.
-In general, knowing a system's position and velocity does not suffice for knowing its momentum.
-Considering a photon as a classical particle, we cannot determine its momentum from its trajectory through space.\footnote{Since photons travel at constant speed, we can recover the direction of momentum---but not the magnitude---from the trajectory.
-Two photons with different momenta can follow the same trajectory in physical space.} Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
-Moreover, different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
+% Argument here could be logically more clear: note that we next have to recover Lagrangian mechanics and do so through an additional condition.
+
+To show that motion equivalence is a further condition, it suffices to show that some Hamiltonian systems violate it.
+This follows from the fact that knowing a system's position and velocity does not always suffice for knowing its momentum.
+Hence, we cannot always determine the evolution of a system in phase space from its trajectory in physical space.
+For instance, if we consider a photon as a classical particle, then we cannot determine its momentum from its trajectory through space.
+Since photons travel at constant speed, we can recover the direction of momentum---but not the magnitude---from the trajectory.
+Two photons with different momenta can follow the same trajectory in physical space.
+Hence, knowledge of the physical trajectory of a photon is insufficient for knowledge of its phase space evolution.
+In general, different Lagrangians---differing even by more than a constant---can correspond to the same particle trajectories while not agreeing on the total energy, even up to a constant.\footnote{\textcites[1185--1186]{Barrett2} discusses this in the context of the failure of equivalence between the vector field formulations of Lagrangian and Hamiltonian mechanics.
 In contrast, the trajectories in phase space (encoded by the Hamiltonian vector field) uniquely define the Hamiltonian up to a constant.} Hence, the set of classical systems that satisfy motion equivalence is a proper subset of the classical systems that satisfy the conditions for Hamiltonian mechanics.
@@
 
+% % To do: address referee 1 concern about the clarity of this argument/example: ``what if momentum is defined in terms of velocity? Or is the idea to know that we also need to know the mass?"
+
 Of course, different definitions of `Hamiltonian' and `Lagrangian' systems can result in different interpretive relationships.
-On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems satisfy the motion equivalence assumption.
-Since Lagrangian systems satisfy all three conditions (infinitesimal reducibility, deterministic/reversible evolution, and motion equivalence), they constitute the intersection of Newtonian and Hamiltonian systems.} We thereby set aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \textcites*[3-4]{Baez}{Earman}{Norton}.} Additionally, we set aside dissipative systems, i.e.
+On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems also satisfy the motion equivalence assumption.
+Since Lagrangian systems satisfy all three conditions (infinitesimal reducibility, deterministic/reversible evolution, and motion equivalence), they lie in the intersection of Newtonian and Hamiltonian systems.} We thereby set aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \textcites*[3-4]{Baez}{Earman}{Norton}.} Additionally, we set aside dissipative systems, i.e.
 those that fail to conserve energy.
@@
 This supports our claim that the property of infinitesimal reducibility fruitfully characterizes classical systems.
-Before presenting the derivational details of our argument in Section~\ref{derivation}, we describe how we avoid central problems facing North's privileging of Hamiltonian mechanics.
-%We view non-dissipative systems as neither Lagrangian nor Hamiltonian. 
-%\footnote{For discussion of the numerous conceptual problems that arise within different approaches to classical mechanics (including point-masses, rigid-bodies, and continua) see \textcites{Wilson}.
-In Wilson's terminology, the framework we apply here may be yet another `theory facade' for classical physics, but a fruitful one nonetheless.}
+Before presenting the details of our argument in Section~\ref{derivation}, we describe how we avoid central problems facing North's privileging of Hamiltonian mechanics.
 
@@
 As indicated at the start of Section~\ref{privileging}, we use `fundamentality' in a non-metaphysical fashion.
-Some differences in fundamentality are differences in generality: more fundamental frameworks require fewer conditions to hold.\footnote{We also embrace related non-metaphysical uses of `fundamental,' discussed in Section~\ref{density}.} Our notion of fundamentality leads to a correspondingly non-metaphysical notion of `structure.' Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
+Some differences in fundamentality are differences in generality: provided a theoretical framework still captures the relevant behavior, it is more fundamental if it requires fewer conditions to hold.
+Our notion of fundamentality leads to a correspondingly non-metaphysical notion of `structure.' Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
 Compared with North's \parencites*[]{North} account of structure, our definition has two advantages.
@@
 As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is `fundamental in the bottom-layer sense' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental in a classical world does not matter for the questions we seek to answer.
-Regardless of its metaphysical status, Hamiltonian mechanics is physically and conceptually more fundamental.
+Regardless of its metaphysical status, Hamiltonian mechanics is conceptually more fundamental.
 Indeed, our argument does not privilege any one of the various mathematical formalizations of Hamiltonian mechanics.
@@
 We argue that (i) invariant phase space densities are physically fundamental and (ii) Hamiltonian mechanics provides the proper setting for representing these densities.
+%is this the actual phase space density vs. density for all possible trajectories?
 
@@
 
-This section justifies the physical and mathematical relationships in Figure~\ref{diagram}.
-Sections~\ref{infinitesimal}-\ref{deterministic} derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution.
+% % try to shorten intro to this section, since parts are repetitive relative to section 2
+
+Our derivation has three main steps.
+Section~\ref{infinitesimal} shows that the phase space of infinitesimally reducible systems must have at least the structure of a symplectic manifold.
+For systems with deterministic and reversible evolution, Section~\ref{deterministic} shows they must evolve according to a symplectomorphism.
+Hence, we derive the formalism of Hamiltonian mechanics from the conditions of infinitesimal reducibility and deterministic/reversible evolution.
 Section~\ref{motion} shows how systems that satisfy the further condition of motion equivalence can be treated within the formalism of Lagrangian mechanics.
-We thus demonstrate in detail our methodology of deriving  relevant mathematical structures from underlying physical conditions.
-
-Before starting, we want to make it clear that the following is not and cannot be a formal proof.
-The idea is that the assumptions put constraints on the physical entities and, in turn, on the mathematical objects that are used to represent them.
-These constraints are constitutive in nature, they provide requirements for the math, and therefore they cannot be derived from the math.
-Each assumption will introduce multiple constraints, which will be introduced gradually within each section.
-The goal is to show how such constraints follow logically from the assumptions, how they are encoded in the math and how they recover the well known structures.
+
+%-\ref{deterministic} derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution.
+along with a few minor assumptions introduced below
+
+Figure~\ref{diagram} summarizes the results of this derivation.
+The left column lists the three central physical conditions.
+Each condition entails physically significant properties, described in the middle column.
+Each physical property corresponds to a formal/mathematical structure depicted in the right column, including phase space, Hamiltonian mechanics, and Lagrangian mechanics.
+In this way, our framework develops a correspondence between physical conditions and their associated mathematical formalisms.
+We thereby elucidate the physical significance of these formalisms.
+%details of this figure will become clear as we go along.
+
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=\textwidth]{Diagram.png}
+\caption{Physical conditions and corresponding properties and formalisms}
+\label{diagram}
+\end{figure}
+
+Admittedly, our derivation falls short of some norms on mathematical rigor.
+We use the three physical conditions to place constraints on physical entities and, in turn, on the mathematical objects that can be used to represent them.
+Each condition leads to additional constraints, introduced gradually within each section.
+Some of these constraints are regulative in nature: they provide requirements for the math and therefore cannot be derived from mathematics alone.
+Our goal is to show how these constraints follow logically from the physical conditions, how they are encoded in the math, and how they recover well-known structures.
+%to do: give a gloss on notion of regulative assumption (Kant and Peirce)
 
@@
 
-First, we will show that a classical phase space arises if a system satisfies the following physical condition:
+First, we will show that a classical phase space arises if a system satisfies the following physical condition 
+%and a series of minor regulative assumptions:
+%should we indicate to the reader that the following assumption alone is insufficient to recover symplectic manifold, but it is the central assumption.
 
@@
 
-Let $\mathcal{C}$ be the state space for the whole system, so that any point $c \in \mathcal{C}$ represents a particular state of the whole system.
+Let $\mathcal{C}$ be the state space for the whole system, so that any $c \in \mathcal{C}$ represents a particular state of the whole system.
 Given infinitesimal reducibility, a classical system consists of a collection of classical particles.
 Let $\mathcal{S}$ be the state space for the particles, so that each point $s \in \mathcal{S}$ describes a possible state of a classical particle.
-Then for each state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
+Then for each system state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
 That is, the state of the whole system is a distribution over its infinitesimal parts.
 This distribution describes the proportion of classical particles in any particular state from $\mathcal{S}$.
+%ref 1: worth clarifying relationship between particle statespace S and system state space C.
+Is there an S_i for each i-th particle?
+And explain why we are not focusing on the statespace of the many particle system, e.g.
+a cartesian product of state spaces for each particle
 
@@
 
-When the system is characterized by continuous quantities, $\mathcal{S}$ is a manifold, and we define $\mu(U) = \int_U \rho_c(s) d\mathcal{S} $.
-State variables $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates' and $\xi$ is a coordinate map.
+When the system is characterized by continuous quantities, $\mathcal{S}$ is a manifold.
+For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
+\textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates' and $\xi$ is a coordinate map.
 We will have multiple manifolds, so to avoid confusion, we will call the coordinates of the state space `state variables'.
@@
 
-Since $\rho_c$: $\mathcal{S} \to \mathbb{R}$ is a function of the particle state $s$, its value is independent of the choice of state variables $\xi^a$ (i.e.
+Since $\rho_c$: $\mathcal{S} \to \mathbb{R}$ is a function of the particle state $s$, we assume that its value is independent of the choice of state variables $\xi^a$ (i.e.
 the choice of state variable map $\xi$).
-Hence, the density $\rho_{c, \xi}$ transforms as a scalar under state variable changes.\footnote{For two state variable maps $\xi$ and $\hat{\xi}$, we have $\rho_c(s) = (\rho_c \circ \xi^{-1} \circ \xi)(s) = (\rho_c \circ \hat{\xi}^{-1} \circ \hat{\xi})(s)$, which entails that $\rho_c(s) = \rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
-Here, $\xi(s) \equiv \xi^a$ and $\hat{\xi}(s) \equiv \hat{\xi}^b$.} Likewise, the integral $\int_U \rho_c(s) d\mathcal{S} $ is independent of the choice of state variables, so $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.
-By the change of variables formula, we also know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\phi \equiv \hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
+If this assumption did not hold, then we could not model the system in the first place using state variables, and this would preclude many standard representational tools.
+Hence, the density $\rho_{c, \xi}$ must transform as a scalar under state variable changes.\footnote{For two state variable maps $\xi$ and $\hat{\xi}$, we have $\rho_c(s) = (\rho_c \circ \xi^{-1} \circ \xi)(s) = (\rho_c \circ \hat{\xi}^{-1} \circ \hat{\xi})(s)$, which entails that $\rho_c(s) = \rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
+Here, $\xi(s) \equiv \xi^a$ and $\hat{\xi}(s) \equiv \hat{\xi}^b$.} Likewise, we assume that the integral $\int_U \rho_c(s) d\mathcal{S} $ is independent of the choice of state variables, so $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.\footnote{Here, we assume that $\xi$ and $\hat{\xi}$ map the background measure on $\mathcal{S}$ to the standard Lebesgue measure on $\mathbb{R}^n$.
+In order to match integrands below, we also assume that $\xi$ and $\hat{\xi}$ agree on their orientation.
+These are regulative assumptions so that we can model the system using coordinates on $\mathbb{R}^n$.
+They amount to assuming that the particle state space $\mathcal{S}$ has a volume form.} By the change of variables formula, we also know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\phi \equiv \hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
 For convenience, we denote $\rho_{c, \hat{\xi}} (\phi(\xi^a))$ by $\rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
@@
 By dimensional homogeneity, it must also be dimensionless.
-We will now show that in order for the density $\rho_{c, \xi}$ to have these transformation properties, the classical phase space must be at least a symplectic manifold.
+
+We will now show that in order for the density $\rho_{c, \xi}$ to have these transformation properties, the particle phase space $\mathcal{S}$ must be at least a symplectic manifold.
+In other words, given that i) we assume we can model a classical system by a density over its infinitesimal states and ii) these transformation properties are required for the density to play this role, it follows that the phase space must have a symplectic manifold structure.
+Our next assumption involves the relationship between state variables and measurement units.
+%Since these transformation properties are required to use a density to model a classical mechanical system, it follows that having the structure of a symplectic manifold is likewise required.  
 %From the preceding two equations, we see that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.
-
+% $d \xi^a$ and $d \hat{\xi}^b$ are the
+
+We assume that a proper subset of the state variables suffices to define a unit system for all of the state variables.
+Again, this assumption is regulative in the sense that without it, we could not quantify properties of the system using measured values.
+Using this assumption, we will later show that the state space $\mathcal{S}$ must be even-dimensional and hence a candidate for having symplectic structure.
 We define a \textit{unit variable} $q \in \xi^a$ as a particular state variable that defines a unit.
@@
 For example, one can measure vertical position in meters while measuring horizontal position in miles.
-In general, the unit variables $q^i$ are the subset of the state variables that define the unit system, which relates the properties of the system to measured values.
-The $q^i$ determine the units not only for themselves, but for all other state variables.
+The \textit{unit variables} $q^i$ are a privileged subset of the state variables that define the unit system.
+These $q^i$ determine the units not only for themselves, but for all other state variables.
 A \textit{change of units} is therefore a map $\hat{q}^j = \hat{q}^j(q^i)$ that defines new unit variables in terms of old unit variables.
 Since the unit variables define the units for all state variables, this change must induce a unique change of state variables $\hat{\xi}^b = \hat{\xi}^b(\xi^a)$ for all variables.
-
-Let us consider the simplest case: suppose that a single unit variable $q $ suffices to define the unit system for the particle state space.
+%to do: cite susan sterrett work on units as a `logical constraint'; , i.e. they fix the units for all state variables. 
+%ref 2: mention that any transformation on the unit variables `is required to extend to a unique transformation of all the coordinates'
+
+Consider first the simplest case: suppose that a single unit variable $q $ suffices to define the unit system for the particle state space.
 Then, the state variables are a set $\xi^a = \{ q, k^\alpha \}$, where \textit{prima facie} the index $\alpha$ could be any non-negative integer.
@@
 
-Extending to the general case, recall that two unit-variables are independent if changing the units for one does not change the units for the other.
+We extend this construction to the general case with multiple unit variables $q^i$.
+Recall that two unit-variables are independent if changing the units for one does not change the units for the other.
 Suppose our particle state space $\mathcal{S}$ is such that its units are fully defined by $n$ independent unit-variables $q^i$.
@@
 We define an \textit{independent degree of freedom} as the space charted by a pair of such variables.
-
-We can now show that $\mathcal{S}$ must have at least the structure of a symplectic manifold.
-Defining an invariant density requires the existence of infinitesimal areas.
-Therefore we need a two-form $\omega$ that assigns an infinitesimal area to each infinitesimal surface.
-This lets us define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$, for $\Sigma \subset \mathcal{S}$ (i.e.
+Each independent degree of freedom thereby defines an infinitesimal area in phase space.
+These play a crucial role in showing that $\mathcal{S}$ must have a non-degenerate, closed two-form $\omega$, i.e.
+a symplectic form.
+
+% %[To do: need to remove `infinitesimal' below and talk in terms of arbitrary surfaces and volumes...; likewise for subsequent paragraphs]
+We now show that a two-form $\omega$ must be well-defined on $\mathcal{S}$.
+Recall that a density is defined in terms of an amount per unit volume.
+Representing the system state $c$ in terms of its infinitesimal parts requires that the density $\rho_{c, \xi} $ is defined on infinitesimal volumes $d \mathcal{S}$ of the phase space.
+Since each independent degree of freedom comprises a pair of state variables, each infinitesimal volume decomposes into independent infinitesimal areas.
+To define the density $\rho_{c, \xi} $, we must be able to measure each of these independent infinitesimal areas.
+This is precisely the role performed by a two-form $\omega$, which assigns an infinitesimal area to each infinitesimal surface.
+We can then define integrals of the form $\int_{\Sigma} \rho_c \omega(d\Sigma)$, for $\Sigma \subset \mathcal{S}$ (i.e.
 the fraction of the system whose parts are in the region $\Sigma$ of state space $\mathcal{S}$).
-Because degrees of freedom are---at least locally---independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
-This means the volume form is given by $\omega^n$.
-Importantly, the volume form cannot be degenerate because otherwise there would be a `volume' in state space that could not contain states.
+Because the degrees of freedom are independent, the total number of states in a volume is the product of the possible configurations of each degree of freedom.
+This means that the volume form is given by $\omega^n$.
+
+Since each volume of phase space contains some possible states, the volume form cannot be degenerate.
+If the volume form were degenerate, then there would be a `volume' in state space that did not contain states.
+But then this region of state space could not be a volume.
+Since the volume form $\omega^n$ is non-degenerate, the two-form $\omega$ must also be non-degenerate.
+Otherwise, there would be infinitesimal surfaces in phase space assigned zero area, i.e.
+there would be states of phase space that $\omega$ does not count.
+But, $\omega$ must count all possible states, in order for the density $\rho_{c, \xi} $ to be well-defined.
+
+ %Importantly, the volume form cannot be degenerate because otherwise there would be a `volume' in state space that could not contain states.
 But such a region of state space would---by definition---not be a volume.
 Therefore the form $\omega$ itself cannot be degenerate (otherwise, some states would not be counted by it).
-% % Invariant densities can only be assigned to infinitesimal areas. Get a citation for this claim
-
-Next, we show that $\omega$ must be closed, using the fact that degrees of freedom are independent.
+%on needing a two-form to measure areas, could possibly cite Strang 1993, ch. 5 or Mac Lane & Birkhoff (1999, Theorem IX.2.2). 
+%note: an infinitesimal area counts the number of possible states in a region of phase space, as given by the two state variables that compose the degree of freedom
+
+Next, we show that $\omega$ must be closed, using the fact that the degrees of freedom are independent.
 If we translate a surface along an independent degree of freedom, this does not change the area of the surface, meaning that the number of states on that surface is invariant.
-Consider a parallelepiped, whose opposite sides have equal areas.
-Integrating over the closed surface of the parallelepiped, opposite sides provide equal and opposite contributions.
-Hence, the integral over the closed surface surrounding this volume must equal zero.
-This shows that the 2-form $\omega$ must be closed.
-Therefore $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold.
+We can therefore construct a parallelepiped $P$, whose opposite sides have equal areas.
+Integrating $\omega$ over the closed surface $\partial P$ of this parallelepiped, opposite sides provide equal and opposite contributions.
+Hence, the integral over this closed surface must equal zero: $\int_{\partial P} \omega =0$.
+By Stokes theorem, the integral of $d \omega$ over this volume $P$ must equal zero as well: $\int_P d \omega = \int_{\partial P} \omega =0$.
+Since, we can approximate any volume $V$ by such parallelepipeds, $\int_V d \omega = 0$ for arbitrary $V$.
+Hence, the integrand $d \omega$ must vanish, entailing that $\omega$ is closed.
+So $\mathcal{S}$ must come equipped with a two-form $\omega$ that is closed and not degenerate, and therefore $\mathcal{S}$ is a symplectic manifold.
+
+ %Consider a parallelepiped, whose opposite sides have equal areas. This shows that the 2-form $\omega$ must be closed.
 
@@
 Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
-Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$ at time $t_0$ uniquely identifies the state of the system.
-Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial state at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e.
+Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$ at time $t_0$ uniquely identifies the state of the system at all times.
+[TO DO (ref 2): provide more motivation/justification for the preservation of the density.
+Does this follow from the preceding, or is it a further (regulative) assumption we have to make?] Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial state at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e.
 the density remains the same throughout the evolution.\footnote{For convenience, we sometimes drop the composite state $c$ and coordinate-choice $\xi$ indices from the invariant density $\rho_{c, \xi} (\xi^a)$.
@@
 %The fact that the density is not coordinate-free is central to our argument in Section~\ref{density}.
+
+% [TO DO: provide additional motivation for why the density does not change during the evolution.
+Why do we assume that the evolution preserves the density?
+How does this relate to Liouville's theorem?
+]
 
@@
 The evolution is therefore a symplectomorphism, corresponding to Hamiltonian evolution.
+[TO DO (ref 2): justify that from a symplectomorphism we have Hamiltonian evolution]
 
@@
 In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.
+[TO DO (ref 1): expand on how our argument relates to standard proofs of Liouville's theorem.
+Is it fair to say that we are privileging Hamiltonian mechanics because it forms the basis for a proof of Liouville's theorem?]
 
@@
 
-The phase space density $\rho$ is physically fundamental because it tracks how much of the system is in each possible state.
+The phase space density $\rho$ is fundamental because it tracks how much of the actual system is in each possible state.
 By representing the distribution, the density quantifies the amount of stuff in each phase space state.
@@
 If we did not define the density, then we would need another account of why this mathematical structure arises for classical systems.}
+%To do: Try to add following clarifying remarks somewhere: in privileging the invariant density, we are not saying that merely possible states are physically fundamental.
+Physical fundamentality presumably deals with actual states.
+However, it is conceptually more general to keep track of all physical possibilities, since the actual world/state is an instance of these more general possibilities.
+We interpret possible worlds/states in the non-metaphysical style of Stalnaker's interpretation of possible worlds.
+Hence, this does not necessarily involve commitment to additional metaphysics.
 
 In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
-If the value of the density at one point is larger than another, then physically there is more matter at this point.
-Likewise, if the value of the density is the same at two different points, this entails that there is the same amount of stuff at those points.
+If the value of the density at one point is larger than another, then physically there is more matter in the state represented by that point.
+Likewise, if the value of the density is the same at two different points, this entails that there is the same amount of stuff in these two states.
 For brevity, we'll refer to this as the property of \textit{point-wise comparability}.

### 2021-08-05 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
ClassMechPhil: directory for new submission; reverted previous to what was submitted

--- 2021-07-12 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
@@
 
-Before starting, we want to make it clear that the following is not and cannot be a formal proof.
-The idea is that the assumptions put constraints on the physical entities and, in turn, on the mathematical objects that are used to represent them.
-These constraints are constitutive in nature, they provide requirements for the math, and therefore they cannot be derived from the math.
-Each assumption will introduce multiple constraints, which will be introduced gradually within each section.
-The goal is to show how such constraints follow logically from the assumptions, how they are encoded in the math and how they recover the well known structures.
-
 \subsection{Infinitesimal reducibility}

### 2021-08-05 Articles/2020-ClassicalMechanicsPhil/Submissions/PoS/classical-phil-article-PoS.tex
ClassMechPhil: directory for new submission; reverted previous to what was submitted

--- 2021-07-26 Articles/2020-ClassicalMechanicsPhil/Submissions/BJPS/classical-phil-article-BJPS.tex
@@
 \usepackage{mathptmx}
+%\usepackage{mathpazo}
 
@@
 
-\DeclareFieldFormat{labeldate}{\mkbibbrackets{#1}} %puts brackets around years in citations
+%\DeclareFieldFormat{labeldate}{\mkbibbrackets{#1}} %puts brackets around years in citations
 
@@
 
-\DeclareFieldFormat{biblabeldate}{\mkbibbrackets{#1}} %puts brackets around years in references
+%\DeclareFieldFormat{biblabeldate}{\mkbibbrackets{#1}} %puts brackets around years in references
 
@@
 \begin{document}
-\title{\vspace{-2em} Hamiltonian Privilege}
-%\author{TBD}
-
-\date{\vspace{-3em} \today}
-
+
+\title{\vspace{-3em} \LARGE Hamiltonian Privilege }			
+%\author{\Large \textsuperscript{*\textdagger}}
+\date{\vspace{-7.9ex} \today}
 \maketitle
-
-% 300 words maximum for BJPS
-\vspace{-3em}
+%
+\vspace{-8mm}
+
+\noindent\rule{\textwidth}{1.25pt}
 \bs
-\begin{abstract}
-\noindent We argue that Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
-Our argument provides a non-metaphysical strategy for privileging one formulation of a theory over another: a formulation is more fundamental at least when it is more general.
-We illustrate this criterion through a novel interpretation of classical mechanics in terms of three physical conditions.
+\noindent {\normalsize We argue that Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
+Our argument provides a non-metaphysical strategy for privileging one formulation of a theory over another: a formulation is more fundamental when it is more general.
+We illustrate this criterion through a novel interpretation of classical mechanics, based on three physical conditions.
 Two of these conditions suffice for recovering Hamiltonian mechanics.
-A third condition is necessary for recovering Lagrangian mechanics.
-This shows that Lagrangian systems are a proper subset of Hamiltonian systems, elucidating the connection between these two formalisms.
-Moreover, we argue that the invariant phase space density is physically fundamental.
-Since Hamiltonian mechanics provides the proper setting for this density, we gain another reason for privileging it over Lagrangian mechanics.
-We interpret Lagrangian mechanics within the Hamiltonian framework, providing a geometric interpretation of the principle of stationary action.
-Finally, we rebut arguments for privileging Lagrangian mechanics, showing where they go wrong.
-\end{abstract}
+A third condition is necessary for Lagrangian mechanics.
+Hence, Lagrangian systems are a proper subset of Hamiltonian systems.
+Finally, we provide a geometric interpretation of the principle of stationary action and rebut arguments for privileging Lagrangian mechanics.
+}
 \es
+\noindent\rule{\textwidth}{1.25pt}
+
+%\vspace{2mm}
 
@@
 The construction at the heart of our argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density.
-We show that the invariant density fits more naturally within the Hamiltonian framework, providing another reason to privilege it over Lagrangian mechanics.
+%We show that the invariant density fits more naturally within the Hamiltonian framework, providing another reason to privilege it over Lagrangian mechanics.
 % Our formalism also provides a novel interpretation of the principle of stationary action within Hamiltonian mechanics.
@@
 This completes our primary argument for privileging Hamiltonian mechanics.
-
-Subsequent sections describe additional advantages of Hamiltonian mechanics, along with further interpretive benefits of our approach.
-Section~\ref{density} explains how the invariant phase space density plays a special role in the Hamiltonian framework.
-The density allows us to compare and track the amount of matter in different possible states.
-In contrast, Lagrangian mechanics does not support these important properties of the density.
-Finally, Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework.
+Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework.
 We provide a geometric interpretation of the action principle, which applies even outside the hyper-regular domain.
 We end by rebutting objections that arise from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
-%Along the way, we examine the gauge redundancy of both the conjugate momentum (Section~\ref{gauge_momentum}) and the Lagrangian (Section~\ref{gauge}).
-
-%Despite being coordinate-invariant, one cannot make the density truly coordinate independent. We therefore argue that some quantities are physically significant despite being coordinate-dependent.
 
@@
 
-In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by `fundamental.' Of various notions of fundamentality, we mean something entirely non-mysterious.
-Given two formalisms that capture a class of systems, if one is logically more general, then it is conceptually more fundamental.\footnote{In Section~\ref{density}, we embrace a related non-metaphysical notion of fundamentality that pertains to physical entities rather than theoretical frameworks.} Conceptual fundamentality is familiar from mathematics.
+In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by ``fundamental.'' Of various notions of fundamentality, we mean something entirely non-mysterious.
+Given two formalisms that capture a class of systems, if one is logically more general, then it is conceptually more fundamental.
+Conceptual fundamentality is familiar from mathematics.
 A more fundamental framework encompasses a less fundamental one as a special case.
-It is in this sense that the generalized Stokes theorem is more fundamental than well-known special cases, such as the fundamental theorem of calculus and Green's theorem.
+It is in this sense that the generalized Stokes' theorem is more fundamental than well-known special cases, such as the fundamental theorem of calculus and Green's theorem.
 Similarly, an axiomatization which includes additional axioms is less fundamental than one which drops these axioms while maintaining the same expressive power.
@@
 We do not mean that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in this sense.
+%\footnote{In Section~\ref{density}, we embrace a related non-metaphysical notion of fundamentality that pertains to physical entities rather than theoretical frameworks.}
 
@@
 Importantly, we require that the theory remains wholly \textit{about} classical mechanics.
-Hamiltonian mechanics is more general than Lagrangian mechanics while still being about classical systems, where `aboutness' can be understood using Yablo's framework \parencites*[]{Yablo}.
+Hamiltonian mechanics is more general than Lagrangian mechanics while still being about classical systems, where ``aboutness'' can be understood using Yablo's framework \parencites*[]{Yablo}.
 Having an account of \textit{aboutness} or \textit{subject matter} is a general philosophical problem that any account of fundamentality must address.
@@
 %These bolster our claim that this is an appropriate sense of fundamentality for philosophers to focus on.
+
+% From Healey 2007, (weak) separability says that ``any physical process occupying space-time region R supervenes upon an assignment of qualitative intrinsic physical properties at space-time points in R (and/or in arbitrarily small neighborhoods of those points)'' [46]
+% Non-separability: ``some physical process occupying a region R of space-time is not supervenient upon an assignment of qualitative intrinsic physical properties at space-time points in R'' [124]
+% Strong non-separability: ``some physical process occupying a region R of space-time is not supervenient upon an assignment of qualitative intrinsic physical properties at points of R and/or in arbitrarily small neighborhoods of those points'' [124].
 
@@
 Infinitesimal reducibility distinguishes classical from quantum systems, where the latter are irreducible (i.e.
-we cannot specify the whole state by specifying the states of its subsystems).
-We show in Section~\ref{infinitesimal} that the phase space of infinitesimally reducible systems must have at least the structure of a symplectic manifold.
+we cannot specify the whole state by specifying the states of its subsystems).\footnote{The conditions of infinitesimal (ir)reducibility are similar to various notions of (non-)separability in the literature on physical holism \parencites[46, 124]{Healey}.
+We adopt our preferred terminology to avoid controversies surrounding how best to characterize separability.} We show in Section~\ref{infinitesimal} that the phase space of infinitesimally reducible systems must have at least the structure of a symplectic manifold.
 Recovering Hamiltonian mechanics then simply requires identifying when trajectories on phase space obey Hamiltonian evolution  (i.e.
@@
 We thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution.
+%Since we won't have to settle controversies surrounding these terms, we adopt our preferred terminology.
 
@@
 As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
-% Argument here could be logically more clear: note that we next have to recover Lagrangian mechanics and do so through an additional condition.
 
@@
 
-% % To do: address referee 1 concern about the clarity of this argument/example: ``what if momentum is defined in terms of velocity? Or is the idea to know that we also need to know the mass?"
-
-Of course, different definitions of `Hamiltonian' and `Lagrangian' systems can result in different interpretive relationships.
+% % optional: address refer 1 concern about the clarity of this argument/example: ``what if momentum is defined in terms of velocity? Or is the idea to know that we also need to know the mass?"
+
+Of course, different definitions of ``Hamiltonian'' and ``Lagrangian'' systems can result in different interpretive relationships.
 On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems also satisfy the motion equivalence assumption.
@@
 Although both the Lagrangian and Hamiltonian formulations apply to dissipative systems, we exclude these.\footnote{For details, see \textcites[\S 10.4]{Cline}.
-Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems, although he seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
-For details of the Lagrangian approach to dissipative systems, see \textcites[]{Smith}.} This restricts our claim about the fundamentality of Hamiltonian mechanics to the non-dissipative regime.
+Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems.
+For details of a Lagrangian approach to dissipative systems, see \textcites[]{Smith}.} This restricts our claim about the fundamentality of Hamiltonian mechanics to the non-dissipative regime.
 Finally, our treatment is not restricted to $n$-particle mechanics but applies as well to rigid-body and continuum mechanics.
@@
 Before presenting the details of our argument in Section~\ref{derivation}, we describe how we avoid central problems facing North's privileging of Hamiltonian mechanics.
+%, although Curiel seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
 
@@
 
-As indicated at the start of Section~\ref{privileging}, we use `fundamentality' in a non-metaphysical fashion.
+As indicated at the start of Section~\ref{privileging}, we use ``fundamentality'' in a non-metaphysical sense.
 Some differences in fundamentality are differences in generality: provided a theoretical framework still captures the relevant behavior, it is more fundamental if it requires fewer conditions to hold.
-Our notion of fundamentality leads to a correspondingly non-metaphysical notion of `structure.' Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
+Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure.'' Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
 Compared with North's \parencites*[]{North} account of structure, our definition has two advantages.
@@
 
-Owing to the epistemic difficulty of accessing these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While we will argue that Hamiltonian mechanics is significant partly because it elucidates invariant densities (Section~\ref{density}), this does not entail that additional mathematical structure is physically insignificant or surplus.
+Owing to the epistemic difficulty of accessing these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While Hamiltonian mechanics is significant partly because it elucidates invariant densities, this does not entail that additional mathematical structure is physically insignificant or surplus.
 Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration.
@@
 
-Additionally, whereas North \parencites*[76]{North} seeks to determine the `real' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure ontologically, but as a formal device for encoding physical information or a calculational tool.
-As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is `fundamental in the bottom-layer sense' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental in a classical world does not matter for the questions we seek to answer.
+Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real'' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure ontologically, but as a formal device for encoding physical information or a calculational tool.
+As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is ``fundamental in the bottom-layer sense'' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental in a classical world does not matter for the questions we seek to answer.
 Regardless of its metaphysical status, Hamiltonian mechanics is conceptually more fundamental.
@@
 Motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.
-Section~\ref{density} provides a further argument for viewing Hamiltonian mechanics as more fundamental.
+%Section~\ref{density} provides a further argument for viewing Hamiltonian mechanics as more fundamental.
 We argue that (i) invariant phase space densities are physically fundamental and (ii) Hamiltonian mechanics provides the proper setting for representing these densities.
@@
 \label{derivation}
-
-% % try to shorten intro to this section, since parts are repetitive relative to section 2
 
@@
 
-%-\ref{deterministic} derive the formalism of Hamiltonian mechanics from the assumptions of infinitesimal reducibility and deterministic/reversible evolution.
-along with a few minor assumptions introduced below
+We use the three physical conditions to place constraints on both physical entities and, in turn, the mathematical objects that can be used to represent them.
+Each condition leads to additional constraints, introduced gradually within each section.
+Some of these constraints are regulative in nature: they provide requirements for the math and therefore cannot be derived from mathematics alone.
+Our goal is to show how these constraints follow logically from the physical conditions, how they are encoded in the math, and how they recover well-known structures.
 
@@
 
-Admittedly, our derivation falls short of some norms on mathematical rigor.
-We use the three physical conditions to place constraints on physical entities and, in turn, on the mathematical objects that can be used to represent them.
-Each condition leads to additional constraints, introduced gradually within each section.
-Some of these constraints are regulative in nature: they provide requirements for the math and therefore cannot be derived from mathematics alone.
-Our goal is to show how these constraints follow logically from the physical conditions, how they are encoded in the math, and how they recover well-known structures.
-%to do: give a gloss on notion of regulative assumption (Kant and Peirce)
+%Like many arguments in physics, our derivation falls short of some norms on mathematical rigor. 
+%optional to do: give a gloss on notion of regulative assumption (Kant and Peirce)
 
@@
 First, we will show that a classical phase space arises if a system satisfies the following physical condition
-%and a series of minor regulative assumptions:
-%should we indicate to the reader that the following assumption alone is insufficient to recover symplectic manifold, but it is the central assumption.
 
@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.}
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.}
 % % % For further discussion, see \textcites{shpmp}. [omit this for review].
@@
 Let $\mathcal{S}$ be the state space for the particles, so that each point $s \in \mathcal{S}$ describes a possible state of a classical particle.
-Then for each system state $c \in \mathcal{C}$ there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
+Then for each system state $c \in \mathcal{C}$, there exists a unique distribution $\rho_c : \mathcal{S} \to \mathbb{R} $ that describes the state of $c$'s parts.
 That is, the state of the whole system is a distribution over its infinitesimal parts.
@@
 For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
-\textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called `coordinates' and $\xi$ is a coordinate map.
-We will have multiple manifolds, so to avoid confusion, we will call the coordinates of the state space `state variables'.
-As usual with coordinate charts, $\xi$ is properly defined on some $U \subseteq \mathcal{S}$, and $\xi^{-1}$ is defined on $\xi(U) \subseteq \mathbb{R}^n$.
+\textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called ``coordinates'' and $\xi$ is a coordinate map.
+To avoid confusion, we will call the coordinates of the state space $\mathcal{S}$ ``state variables.'' As usual with coordinate charts, $\xi$ is properly defined on some $U \subseteq \mathcal{S}$, and $\xi^{-1}$ is defined on $\xi(U) \subseteq \mathbb{R}^n$.
 We will neglect these subtleties in what follows.} The map $\xi$ represents a state $s \in \mathcal{S}$ by a set of state variables $\xi^a \in \mathbb{R}^n$, such that $\xi (s) = \xi^a$.
@@
 This density plays a crucial role throughout our argument.
+It allows us to compare and track the amount of matter in different possible states.
 
@@
 Hence, the density $\rho_{c, \xi}$ must transform as a scalar under state variable changes.\footnote{For two state variable maps $\xi$ and $\hat{\xi}$, we have $\rho_c(s) = (\rho_c \circ \xi^{-1} \circ \xi)(s) = (\rho_c \circ \hat{\xi}^{-1} \circ \hat{\xi})(s)$, which entails that $\rho_c(s) = \rho_{c, \xi} (\xi^a) = \rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
-Here, $\xi(s) \equiv \xi^a$ and $\hat{\xi}(s) \equiv \hat{\xi}^b$.} Likewise, we assume that the integral $\int_U \rho_c(s) d\mathcal{S} $ is independent of the choice of state variables, so $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.\footnote{Here, we assume that $\xi$ and $\hat{\xi}$ map the background measure on $\mathcal{S}$ to the standard Lebesgue measure on $\mathbb{R}^n$.
+Here, $\xi(s) \equiv \xi^a$ and $\hat{\xi}(s) \equiv \hat{\xi}^b$.} Likewise, we assume that the integral $\int_U \rho_c(s) d\mathcal{S} $ is independent of the choice of state variables, so that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b$.\footnote{Here, we assume that $\xi$ and $\hat{\xi}$ map the background measure on $\mathcal{S}$ to the standard Lebesgue measure on $\mathbb{R}^n$.
 In order to match integrands below, we also assume that $\xi$ and $\hat{\xi}$ agree on their orientation.
 These are regulative assumptions so that we can model the system using coordinates on $\mathbb{R}^n$.
-They amount to assuming that the particle state space $\mathcal{S}$ has a volume form.} By the change of variables formula, we also know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\phi \equiv \hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
+They amount to assuming that the particle state space $\mathcal{S}$ has a volume form.} Applying the change of variables formula, we know that $\int_{\hat{\xi}(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) d \hat{\xi}^b= \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.\footnote{The change of variables transformation from $\xi^a$ to $\hat{\xi}^b$ is given by $\phi \equiv \hat{\xi} \circ \xi^{-1}$: $\mathbb{R}^n \to \mathbb{R}^n $.
 For convenience, we denote $\rho_{c, \hat{\xi}} (\phi(\xi^a))$ by $\rho_{c, \hat{\xi}} (\hat{\xi}^b)$.
@@
 
+\subsubsection{The particle state space is even-dimensional}
+
 We assume that a proper subset of the state variables suffices to define a unit system for all of the state variables.
 Again, this assumption is regulative in the sense that without it, we could not quantify properties of the system using measured values.
-Using this assumption, we will later show that the state space $\mathcal{S}$ must be even-dimensional and hence a candidate for having symplectic structure.
+Using this assumption, we will show that the state space $\mathcal{S}$ must be even-dimensional and hence a candidate for having symplectic structure.
+
 We define a \textit{unit variable} $q \in \xi^a$ as a particular state variable that defines a unit.
@@
 We define an \textit{independent degree of freedom} as the space charted by a pair of such variables.
-Each independent degree of freedom thereby defines an infinitesimal area in phase space.
+Each independent degree of freedom thereby defines a two-dimensional surface in phase space.
 These play a crucial role in showing that $\mathcal{S}$ must have a non-degenerate, closed two-form $\omega$, i.e.
@@
 
-% %[To do: need to remove `infinitesimal' below and talk in terms of arbitrary surfaces and volumes...; likewise for subsequent paragraphs]
-We now show that a two-form $\omega$ must be well-defined on $\mathcal{S}$.
+\subsubsection{The particle state space has a symplectic form}
+
+We now show that $\mathcal{S}$ must have a well-defined two-form $\omega$.
 Recall that a density is defined in terms of an amount per unit volume.
@@
 Since each volume of phase space contains some possible states, the volume form cannot be degenerate.
-If the volume form were degenerate, then there would be a `volume' in state space that did not contain states.
+If the volume form were degenerate, then there would be a ``volume'' in state space that did not contain states.
 But then this region of state space could not be a volume.
@@
 there would be states of phase space that $\omega$ does not count.
-But, $\omega$ must count all possible states, in order for the density $\rho_{c, \xi} $ to be well-defined.
-
- %Importantly, the volume form cannot be degenerate because otherwise there would be a `volume' in state space that could not contain states.
-But such a region of state space would---by definition---not be a volume.
-Therefore the form $\omega$ itself cannot be degenerate (otherwise, some states would not be counted by it).
+But, $\omega$ must count all possible states in order for the density $\rho_{c, \xi} $ to be well-defined.
+
 %on needing a two-form to measure areas, could possibly cite Strang 1993, ch. 5 or Mac Lane & Birkhoff (1999, Theorem IX.2.2). 
@@
 Hence, the integral over this closed surface must equal zero: $\int_{\partial P} \omega =0$.
-By Stokes theorem, the integral of $d \omega$ over this volume $P$ must equal zero as well: $\int_P d \omega = \int_{\partial P} \omega =0$.
+By Stokes' theorem, the integral of $d \omega$ over this volume $P$ must equal zero as well: $\int_P d \omega = \int_{\partial P} \omega =0$.
 Since, we can approximate any volume $V$ by such parallelepipeds, $\int_V d \omega = 0$ for arbitrary $V$.
@@
 
- %Consider a parallelepiped, whose opposite sides have equal areas. This shows that the 2-form $\omega$ must be closed.
-
-The units we choose for $\omega$ are really the units we choose to count states over an independent degree of freedom.
+Moreover, the units we choose for $\omega$ are the units we choose to count states over an independent degree of freedom.
 This agrees with the practice in statistical mechanics.
@@
 
-The conclusion is that, under the infinitesimal reducibility assumption, a state $c$ of the system is represented by a coordinate-invariant density over the state space $\mathcal{S}$ of its infinitesimal parts.
+In sum, under the infinitesimal reducibility assumption, a state $c$ of the system is represented by a coordinate-invariant density over the state space $\mathcal{S}$ of its infinitesimal parts.
 We have shown that the state space of the particles must be a symplectic manifold.
@@
 Therefore symplectic structure is exactly the structure needed to define an invariant density.
-Section~\ref{density} describes further interpretive consequences of this connection between invariant densities and symplectic structure.
-% % [XXX: clarify/clean-up this ending.
-also, emphasize/foreshadow the physical signifiance of the distribution and density].
-also, clarify relationship b/w distribution (ontology) and density (epistemic access to/representation of this ontology)
-%Symplectic manifold structure is exactly the structure needed to define an invariant density.
+
+%Section~\ref{density} describes further interpretive consequences of this connection between invariant densities and symplectic structure.
 
@@
 
-In our case, this applies to both the state of the whole system and the state of its parts.
-Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
-Assuming deterministic and reversible evolution, the initial state $s_0 = \lambda(t_0)$ at time $t_0$ uniquely identifies the state of the system at all times.
-[TO DO (ref 2): provide more motivation/justification for the preservation of the density.
-Does this follow from the preceding, or is it a further (regulative) assumption we have to make?] Moreover, if $\rho(\lambda(t_0), t_0)$ is the density associated to the initial state at the initial time, we must have that $\rho(\lambda(t), t) = \rho(\lambda(t_0), t_0)$, i.e.
-the density remains the same throughout the evolution.\footnote{For convenience, we sometimes drop the composite state $c$ and coordinate-choice $\xi$ indices from the invariant density $\rho_{c, \xi} (\xi^a)$.
-We reserve the word `distribution' for the coordinate-free distribution $\rho_c (s)$.
-} In other words, all particles that begin in a given initial state must end up in the same final state and vice-versa.
-%The fact that the density is not coordinate-free is central to our argument in Section~\ref{density}.
-
-% [TO DO: provide additional motivation for why the density does not change during the evolution.
-Why do we assume that the evolution preserves the density?
-How does this relate to Liouville's theorem?
-]
-
-Now consider the integral $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
+\noindent In our case, this applies to both the state of the whole system and the state of its parts.
+
+First, we will show that deterministic and reversible evolution entails that the density $\rho_{c, \xi}$ is constant in time.
+We consider a particle state $s_0$ at time $t_0$.
+Given deterministic and reversible evolution, there exists a unique evolution $\lambda: \mathbb{R} \to \mathcal{S}$ such that $\lambda(t_0) = s_0$.
+Hence, the initial state $s_0$ at time $t_0$ uniquely identifies the state of this particle at all times.
+Let $\rho(s, t)$ be the density associated to a state $s$ at a time $t$.\footnote{For convenience, we sometimes drop the composite state $c$ and coordinate-choice $\xi$ indices from the invariant density $\rho_{c, \xi} (\xi^a)$.
+We reserve the word ``distribution'' for the coordinate-free distribution $\rho_c (s)$.
+} Then  $\rho(\lambda(t_0), t_0)$ represents the part of the composite system in particle state $s$ at that time.
+Likewise, for $t_1 > t_0$,  $\rho(\lambda(t_1), t_1)$ represents the part of the system associated to the evolved state at this future time.
+Given deterministic and reversible evolution, all particles that start in a given initial state must end up in the same final state, and vice-versa.
+Hence, $\rho(\lambda(t_0), t_0) = \rho(\lambda(t_1), t_1)$ for all $t_0,\, t_1$.
+In other words, the density remains the same throughout the evolution.
+%Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
+
+Second, we show that the two-form $\omega$ is invariant under this evolution, and hence that the system follows Hamiltonian evolution.
+Consider a surface $\Sigma_0 \subset \mathcal{S}$ of particle states at some time $t_0$.
+If two-dimensional, this surface defines an independent degree of freedom.
+The integral $\int_{\Sigma_0} \rho_c (s, t_0) \omega(d\Sigma_0)$ represents the fraction of the system in states defined on $\Sigma_0$.
+Since each particle state evolves by $\lambda$, $\Sigma_0$ evolves to a surface $\Sigma_1$ at time $t_1$.
+This leads to a corresponding $\int_{\Sigma_1} \rho_c (s, t_1) \omega(d\Sigma_1)$, representing the fraction of the system on this evolved surface.
+Since the evolution is deterministic and reversible, all and only the particle states that began on surface  $\Sigma_0$ evolve to $\Sigma_1$.
+Consequently, the fraction of the system on these two surfaces at these two times must be the same: $\int_{\Sigma_0} \rho_c (s, t_0) \omega(d\Sigma_0) = \int_{\Sigma_1} \rho_c (s, t_1) \omega(d\Sigma_1)$.
+Since the density is constant, we know that $\rho_c (s, t_0) = \rho_c (s, t_1)$.
+Hence, the two-form $\omega$ must remain the same.
+That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (otherwise, volumes would not be mapped to equal volumes, violating determinism).
+The evolution is therefore a symplectomorphism and hence corresponds to a Hamiltonian evolution.
+%Now consider the integral $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
 Mapping the region $\Sigma$ to a new region, the fraction of the system found in both the old and new regions must be the same.
 Since both the integral and $\rho_c$ must remain constant during the evolution, the two-form $\omega$ must also remain the same.
-That is, the areas in phase space must be mapped to areas of equal size, and independent degrees of freedom must be mapped to independent degrees of freedom (otherwise, volumes would not be mapped to equal volumes, violating determinism).
-The evolution is therefore a symplectomorphism, corresponding to Hamiltonian evolution.
-[TO DO (ref 2): justify that from a symplectomorphism we have Hamiltonian evolution]
-
-Intuitively, this argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes (i.e.
-sets of states are mapped to sets of states of equal size), which leads to Hamiltonian evolution.
+
+Intuitively, our argument is the inverse of Liouville's theorem: instead of positing Hamiltonian evolution and finding conservation of areas and volumes, deterministic and reversible evolution imposes the conservation of areas and volumes (i.e.
+sets of states are mapped to sets of states of equal size).
+This leads to Hamiltonian evolution.
 In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.
-[TO DO (ref 1): expand on how our argument relates to standard proofs of Liouville's theorem.
+% [(ref 1): expand on how our argument relates to standard proofs of Liouville's theorem.
 Is it fair to say that we are privileging Hamiltonian mechanics because it forms the basis for a proof of Liouville's theorem?]
@@
 \bs \noindent
-\textit{Motion equivalence assumption}: the motion of the system (i.e.
-trajectories in physical space-time) suffices to recover its dynamics (i.e.
+\textit{Motion equivalence}: the motion of the system (i.e.
+trajectories in physical space) suffices to recover its dynamics (i.e.
 evolutions in state space) and vice-versa.\footnote{ \textcites[218]{Abraham}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time; see \textcites[1180-1181]{Barrett2} for discussion.
@@
 
-If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.
+\noindent If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.
+
 Note that each space variable $x^i$ is a unit-variable, i.e.
@@
 It follows that (i) $x^i = q^i$, (ii) each $q^i$ is paired with a conjugate $p_i$, and (iii) each state $\{q^i, p_i\}$ corresponds to one and only one trajectory in physical space.
+
 At each point $x^i$, then, there must be infinitely many trajectories, one for each possible combination of $\{p_i\}$.
@@
 \begin{equation}
-v^i = d_t q^i = \frac{\partial H }{\partial p_i} = \frac{1}{m} g^{ij}(p_j - A_j) \\
+v^i = d_t q^i = \frac{\partial H }{\partial p_i} = \frac{1}{m} g^{ij}(p_j - A_j) 
 \end{equation}
@@
 \begin{equation} \label{massive_Hamiltonian}
-H = \frac{1}{2m} (p_i - A_i) g^{ij}(p_j - A_j) + V \\
+H = \frac{1}{2m} (p_i - A_i) g^{ij}(p_j - A_j) + V 
 \end{equation}
@@
 
-This completes our central argument for privileging Hamiltonian mechanics over Lagrangian mechanics.
+This completes our primary argument for privileging Hamiltonian mechanics over Lagrangian mechanics.
 We have seen that Lagrangian systems are a proper subset of Hamiltonian systems, specifically those Hamiltonian systems that also satisfy the motion equivalence condition.
-In the remainder of this essay, we provide further reasons for privileging Hamiltonian mechanics.
+In the next section, we provide a geometric interpretation of the principle of stationary action within Hamiltonian mechanics.
+Our construction makes manifest the gauge-dependency of the Lagrangian, providing a further reason to privilege the Hamiltonian.
+Finally, we respond to Curiel's arguments against the fundamentality of Hamiltonian mechanics.
+% In the remainder of this essay, we provide further reasons for privileging Hamiltonian mechanics.
 The next section describes how Hamiltonian mechanics provides the natural setting for the invariant density, which characterizes how a classical object's parts are distributed.
 The invariant phase space density is more fundamental than the state variables, since conjugate momentum is a gauge-dependent quantity.
-In Section~\ref{Lagrangian} we provide a geometric interpretation of the principle of stationary action within Hamiltonian mechanics.
-Our construction makes manifest the gauge-dependency of the Lagrangian, providing a further reason to privilege the Hamiltonian.
-Finally, we respond to Curiel's arguments against the fundamentality of Hamiltonian mechanics.
-
-\section{The physical significance of invariant densities}
-\label{density}
-
-Hamiltonian mechanics is more fundamental than Lagrangian mechanics because Hamiltonian systems require fewer physical conditions.
-While this provides one reason to privilege Hamiltonian mechanics, it is not the only one.
-In this section, we argue that the special relationship between Hamiltonian mechanics and invariant densities provides a further reason to privilege Hamiltonian mechanics.
-Additionally, we argue that viewing physics through the lens of invariant densities provides a better understanding of classical systems.\footnote{It also provides a better understanding of statistical mechanics and entropy, since phase space provides a uniform measure over q and p.
-We discuss this elsewhere.}
-% % % For discussion, see \textcites{shpmp} [omit for review].}
-
-The phase space density $\rho$ is fundamental because it tracks how much of the actual system is in each possible state.
-By representing the distribution, the density quantifies the amount of stuff in each phase space state.
-Typically, the canonical position and momentum play the role of specifying a classical state, since they fix the values of all other observables.
-Ruetsche refers to this as ``fundamental in the physicist's sense'' \parencite*[31, 200]{Ruetsche}.
-However, we will show in Section~\ref{gauge_momentum} that the conjugate momentum is gauge-dependent.
-For this reason, the invariant density provides a more fundamental specification of the state than the state variables themselves.
-%\footnote{Section~\ref{infinitesimal} provides another reason for viewing the density as fundamental: by defining the density $\rho_{c, \xi}$, we recover the right mathematical structure, namely a symplectic manifold.
-If we did not define the density, then we would need another account of why this mathematical structure arises for classical systems.}
-%To do: Try to add following clarifying remarks somewhere: in privileging the invariant density, we are not saying that merely possible states are physically fundamental.
-Physical fundamentality presumably deals with actual states.
-However, it is conceptually more general to keep track of all physical possibilities, since the actual world/state is an instance of these more general possibilities.
-We interpret possible worlds/states in the non-metaphysical style of Stalnaker's interpretation of possible worlds.
-Hence, this does not necessarily involve commitment to additional metaphysics.
-
-In the Hamiltonian formalism, the density has the following desirable property: we can compare the density point-wise across phase space.
-If the value of the density at one point is larger than another, then physically there is more matter in the state represented by that point.
-Likewise, if the value of the density is the same at two different points, this entails that there is the same amount of stuff in these two states.
-For brevity, we'll refer to this as the property of \textit{point-wise comparability}.
-Point-wise comparability requires a coordinate-invariant density.
-As the derivation in Section~\ref{infinitesimal} shows, to guarantee the existence of an invariant density, we must define state space over a symplectic manifold.
-In other words, to generically have a point-wise comparable density, we must use canonical coordinates $(q, p)$ defined on a symplectic manifold.
-
-Section~\ref{dependence} shows why point-wise comparability does not hold in general: the density is a coordinate-dependent quantity, and thus it is not generically coordinate-invariant.
-Within the Lagrangian formalism, point-wise comparability holds only when we have made a special choice of coordinates.
-This gives a further reason to privilege Hamiltonian mechanics.
-Section~\ref{gauge_momentum} shows why the conjugate momentum $p$ is gauge-dependent.
-This bolsters our claim that the density is more fundamental than the choice of state variable pairs.
-It also shows that---\textit{pace} conventional wisdom---position and conjugate momentum are \textit{not} on equal physical footing within Hamiltonian mechanics.
-The syntactic symmetry between position and momentum in the mathematical formalism does not entail that these quantities are equally physically significant.
-
-\subsection{Point-wise comparability}
-\label{dependence}
-
-To understand what makes densities special in the context of Hamiltonian mechanics, we first consider densities in general.
-By a `density' we mean the intuitive physical notion of a quantity divided by the region it occupies, such as mass per volume or charge per area.
-Over a continuum, this is an integrable function $\rho(x)$.
-Mathematically, the concept is captured by the Radon–Nikodym derivative: we have (i) a region $U$, (ii) two measures $m$ and $V$, and (iii) the derivative satistifies $m(U) = \int_U \rho dV$.
-This relationship shows that the units of $\rho$ depend on the units of both $m$ and $V$.
-If $\rho(x, y)$ is a function of multiple variables, dimensional analysis requires that the units of $\rho$ are consistent with the units of the arguments in the expression $\int_U \rho(x, y) dx dy$.
-Yet in general, a manifold does not provide a way to measure areas and volumes.
-For example, a manifold may capture the physical characteristics of a group of people in terms of their height $h$ and their weight $w$, leading to a density $\rho(h,w)$ with units of people/(m$\cdot$Kg).
-Or we may represent these characteristics using height $h$ and body mass index $bmi = w / h^2$, which leads to a different density $\rho(h,bmi)$ with units of (people$\cdot$m)/Kg.
-
-Hence, the density is coordinate-dependent because its definition requires us to make a choice of unit-variables.
-Worse than that, it is not generically coordinate-invariant, since the density changes with the Jacobian.
-In the above example, $\rho(h,w) = |J| \rho(h, bmi) = \frac{1}{h^2} \rho(h, bmi)$.
-Similarly, in changing from polar to Cartesian coordinates we have $\rho(x,y) = |J| \rho(r, \theta) = r \rho(r, \theta)$.
-Since the Jacobian is a function of the point, the value of $|J|$ can change from point to point.
-Consequently, the density is not guaranteed to be directly comparable at different points: point-wise comparability need not hold in an arbitrary coordinate system.
-If we have a uniform, constant density in $x$ and $y$, we will have an $r$-dependence in $r$ and $\theta$.
-This dependence is introduced precisely because the region identified by $dr$ and $d\theta$ is $dr \,r d\theta$, which is a function of $r$.
-This makes the value of the density, in general, not directly comparable at different points.
-Without additional structure, there is no objective sense in which the density is higher at one point or another: it depends on the coordinates.
-
-%The Hamiltonian formalism circumvents this further epistemic requirement by defining the density using quantities that vary covariantly.
-Hence, we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in.
-%when we define the density originally, it is coordinate-free (rho as a function of the state space of the particles, S).
-But when we go to express it numerically, then it becomes coordinate-dependent.
-(I.e.
-when we express it over a manifold, then it needs to be a density for integration, and hence it becomes coordinate-dependent).
-
-The Hamiltonian formalism fixes this issue by assigning areas to regions in a way that is independent of the choice of unit-variables.
-To do so, it forces us to restrict ourselves to canonical coordinates, which explains why momentum has to change covariantly.
-The result is that we can perform point-wise comparisons in the Hamiltonian formalism without knowing what coordinate system we are in (in the sense of space-time coordinates).
-Any and all (time-independent) canonical coordinate systems support point-wise density comparisons.\footnote{Time-dependent coordinate transformations require working in the extended phase space $(q^i, p_i, t, E)$.
-See \textcites[]{Struckmeier}.} In contrast, in the Lagrangian framework, there are at best special coordinate systems that support point-wise comparability.\footnote{One might wonder if it is possible to construct a coordinate-independent density, perhaps using either (i) volume forms from differential geometry or (ii) the canonical volume form of Riemannian geometry.
-Both of these strategies fail, although space precludes a demonstration.}
-
-% % Try to precisely gloss/specify the condition for a coordinate system to be special in this way, such that it supports point--wise comparison.
-(Seems inappropriate to mention the Jacobian, since this matters for transforming from one coordinate system to another.)
-
-\subsection{Gauge redundancy of conjugate momentum}
-\label{gauge_momentum}
-
-Whereas we privilege the invariant phase space density, some philosophers have instead privileged the canonical state variables $(q, p)$.
-For instance, in her argument for the fundamentality of Hamiltonian mechanics, North claims that ``momentum is a truly fundamental quantity, on a par with position'' \parencites*[77]{North}.
-Likewise, \textcites[164,171]{Wallace} claims that conjugate momentum and position are on equal footing within Hamiltonian mechanics (but not in general).
-Yet, as we now show, the conjugate momentum $p$ is actually gauge-dependent, even in the context of a free particle.
-Assuming that gauge-invariance is a necessary condition for fundamental physical significance, this precludes $p$ from having fundamental status.
-Instead, we should understand the conjugate momentum $p$ as a label that we can use to express the invariant density over states.
-In this way, the invariant density is more fundamental than the state variables themselves.
-% % after Double checking Wallace citation, I should probably find a better/more direct one. e.g. Earman article in that volume?
-
-We begin with a precise account of what gauge-dependency physically represents.
-As before, we consider a physically-motivated definition that leads to the correct mathematical structure.
-When motion equivalence holds, we can completely characterize states in terms of the position $q$ and its first derivative $\dot{q}$.
-Hence, any canonical transformation in phase space that leaves $(q,\dot{q}) $ invariant \textit{ipso facto} leaves the physical state invariant.
-We will show that these transformations are exactly the gauge transformations.
-Since the position does not change, the new state variables can be written as $(q, \hat{p})$.
-They must respect the usual Poisson brackets because the transformation is canonical:
-\begin{equation}
-\begin{aligned}
-	\{q^i, q^j\} &= 0 \\
-	\{q^i, \hat{p}_j\} &= \delta^i_j \\
-	\{\hat{p}_i, \hat{p}_j\} &= 0
-\end{aligned}
-\end{equation}
-Using the second equation we find:
-\begin{equation}
-\begin{aligned}
-	\{q^i, \hat{p}_j\} &= \frac{\partial q^i}{\partial q^k} \frac{\partial \hat{p}_j}{\partial p_k} - \frac{\partial q^i}{\partial p_k} \frac{\partial \hat{p}_j}{\partial q^k} \\
-	&= \delta^i_k \frac{\partial \hat{p}_j}{\partial p_k} - 0 \frac{\partial \hat{p}_j}{\partial q^k} \Rightarrow \frac{\partial \hat{p}_j}{\partial p_i} = \delta^i_j
-\end{aligned}
-\end{equation}
-Integrating this last equation yields
-\begin{equation}
-\begin{aligned}
-\hat{p}_j = p_j + A_j(q^i)
-\end{aligned}
-\end{equation}
-where the $A_j$ are arbitrary functions of position.
-
-Using the third Poisson bracket, we have:
-\begin{equation}
-\begin{aligned}
-0 = \{\hat{p}_i, \hat{p}_j\} &= \{p_i + A_i, p_j + A_j\} \\
-&= \{p_i, p_j \} + \{p_i , A_j\} + \{A_i, p_j \} + \{A_i, A_j\} \\
-&= 0 - \{A_j, p_i\} + \{A_i, p_j \} + 0 \\
-&= - \frac{\partial A_j}{\partial q^i} + \frac{\partial A_i}{\partial q^j} \\
-&= curl(A_j) \\
-\end{aligned}
-\end{equation}
-Since $A_j$ is curl free, we can express it as the gradient of a function $f$.
-Therefore, changes of state variable of the form $\hat{p}_j = p_j + \partial_j f$ do not change either the equations of motion or the trajectories.
-We recognize these as the usual family of gauge transformations.
-We can thereby claim that, in the context of classical particle mechanics, the gauge transformations are the state variable transformations that leave the density and $(q,\dot{q}) $ invariant.
-
-The gauge-dependency of the conjugate momentum supplies further grounds for privileging the invariant phase space density.
-Insofar as coordinate-invariance is a plausible necessary condition for fundamental physical status, so is gauge-invariance.
-By being gauge-dependent, the conjugate momentum cannot be physically fundamental.
-Nevertheless, it remains physically significant, similar to how coordinate-variant quantities can be physically significant but not fundamental.
-%might not need this paragraph; tacitly references the section 5.2 I deleted.
-
-Additionally, the gauge-dependency of conjugate momentum shows that all state variables are \textit{not} on equal physical footing.
-Although the canonical position and momentum variables are on equal syntactical footing within symplectic mechanics, this does not entail that they are equally physically significant.
-To see this, recall that the phase space density remains invariant upon changing the unit-variables $q$, inducing a state variable transformation on the conjugate momenta $p$ (which compensate by transforming covariantly).
-It is in this sense that the unit-variables $q$---rather than the conjugate momenta $p$---define the units of physical measurement, allowing us to assign numbers to physical quantities.
-Consequently, although symplectic mechanics allows arbitrary mixings of $q$'s and $p$'s that preserve the symplectic manifold, not all of these combinations are equally significant.
-In general, the canonical position coordinates $q$ are epistemically privileged because they are linked with measurements, in a sense that the logically-derivative $p$'s are not.
-Unlike both velocity and the kinetic momentum $m g_{ji} v^i$, conjugate momentum is not an experimentally measurable quantity, because its specification involves a gauge choice.
-%\footnote{As discussed in Section~\ref{fundamental}, these units are physically significant despite not being coordinate-invariant.}
 
@@
 
-By analyzing classical systems in terms of their invariant phase space density, we gain new insights into Lagrangian mechanics.
-Section~\ref{action} interprets the action principle geometrically in extended phase space as a consequence of the conserved density.
-We show that the action principle holds even when the action is not defined as a function of position and velocity, since it can always be expressed as a function of position and conjugate momentum.
+By viewing classical systems through their invariant phase space density, we gain new insights into Lagrangian mechanics.
+Section~\ref{action} uses the invariant phase space density to give a geometric interpretation the action principle.
+We show that this principle holds even when the action is not defined as a function of position and velocity, since it can always be expressed as a function of position and conjugate momentum.
 Our construction makes the gauge-dependence of the Lagrangian manifest.
@@
 the standard phase space extended by the time variable.
-Knowing that the system is deterministic and reversible, it follows that the density of states is invariant.
+Knowing that the system is deterministic and reversible, it follows from Section~\ref{deterministic} that the density of states is invariant.
 This means that possible trajectories of the system in the extended phase space are never pushed toward each other or pulled apart over time.
@@
 This field is divergenceless, $\nabla \cdot \vector{S} = 0$, because for any region $U$, the same number of states will enter the region as exit.
-In particular, note that the displacement is tangent to the trajectory $\gamma [q, p, t] (t)$ in extended phase space, where the variable $t$ functions as both a coordinate and the evolution parameter.
+Moreover, note that the displacement vector $\vec{S}$ is tangent to the trajectory $\gamma [q, p, t] (t)$ in extended phase space, where the variable $t$ functions as both a coordinate and the evolution parameter.
 
@@
 
-Treating time $t$ as both a coordinate and the evolution parameter, the time component of the displacement $S_t = dt/ dt = 1$.\footnote{In the expression `$dt/dt$', the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
+Treating time $t$ as both a coordinate and the evolution parameter, the time component of the displacement $S_t = dt/ dt = 1$.\footnote{In the expression ``$dt/dt$,'' the time variable in the numerator functions as a coordinate, while the time variable in the denominator functions as a parameter.
 We could instead use a parameter $\tau$, and the derivation would remain valid.} This places another constraint on $\vector{\theta} $, since $\vector{S} = -curl(\vector{\theta}) $.
@@
 To derive Hamilton's principle of stationary action, it suffices to show that $\delta \mathscr{A} [\gamma] =\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = 0$.
-%\footnote{When motion equivalence holds, the Legendre transform is well-defined, allowing us to express $L_p$ as a function of $(q, \dot{q}, t)$, i.e. $L_p = L(q, \dot{q}, t) = p \dot{q} - H$.}
 
@@
 
-Curiel argues that Lagrangian mechanics represents classical mechanical systems in a \textit{natural} manner whereas Hamiltonian mechanics involves an essentially artificial construction \parencites*[270]{Curiel}.\footnote{Curiel characterizes `natural' in terms of natural coordinate systems, defined using the almost-tangent structure of the tangent bundle \parencites*[290-291]{Curiel}.
+Curiel argues that Lagrangian mechanics represents classical mechanical systems in a \textit{natural} manner whereas Hamiltonian mechanics involves an essentially artificial construction \parencites*[270]{Curiel}.\footnote{Curiel characterizes ``natural'' in terms of natural coordinate systems, defined using the almost-tangent structure of the tangent bundle \parencites*[290-291]{Curiel}.
 He also uses the almost-tangent structure to give a general characterization of what he takes to be the fundamental kinematical constraints.} According to Curiel, all physically-meaningful classical systems satisfy a certain set of kinematic constraints.
-In the simplest case, these constraints characterize velocity as the first temporal derivative of position: $v^i = d_t x^i = v^i(q^j, p_k)$ (Equation~\ref{weak_equivalence}).\footnote{Curiel treats this constraint as `kinematical' in the sense that it must hold independently of the interactions between a system and its environment.
+In the simplest case, these constraints characterize velocity as the first temporal derivative of position: $v^i = d_t x^i = v^i(q^j, p_k)$ (Equation~\ref{weak_equivalence}).\footnote{Curiel treats this constraint as ``kinematical'' in the sense that it must hold independently of the interactions between a system and its environment.
 He interprets it as a substantive physical condition, rather than an analytic truth \parencites*[282]{Curiel}.} Curiel argues that Lagrangian mechanics encodes the necessary and sufficient conditions for a system to satisfy these kinematic constraints, both within and outside the hyper-regular domain \parencites*[307--308, 311]{Curiel}.
@@
 There, we saw how both the linear relationship between momentum and velocity and the quadratic form (\ref{6.2}) follow from imposing the motion equivalence condition.
-In Equation~\ref{linear_relationship}, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function (connected to the gauge-dependency of the conjugate momentum).
+In Equation~\ref{linear_relationship}, we recovered the linear relationship between conjugate momentum and velocity, up to the addition of an arbitrary function.
 Just as Curiel does, we then used this relationship to derive a quadratic form for the Hamiltonian (Equation~\ref{massive_Hamiltonian}).
@@
 Instead, it arises in all contexts where motion equivalence holds, i.e.
-where the trajectory of the system in physical space-time is necessary and sufficient for its evolution in state space.
+where the trajectory of the system in physical space is necessary and sufficient for its evolution in state space.
 Thus, we take ourselves to have responded to Curiel's main criticism of Hamiltonian mechanics, namely that it must impose \textit{ad hoc} the condition (\ref{6.2}).
@@
 
-Furthermore, Curiel's definition of `classical systems' leads to interpretive challenges that our framework avoids.
+Furthermore, Curiel's definition of ``classical systems'' leads to interpretive challenges that our framework avoids.
 His prohibition against Hamiltonians that violate (\ref{6.2}) excludes Hamiltonians for non-hyper-regular systems.
@@
 For example, consider the case where the Lagrangian $L$ equals zero.
-For this `system,' all paths have the same action, so all paths are possible.
+For this ``system,'' all paths have the same action, so all paths are possible.
 Given an initial probability distribution over states, the Lagrangian fails to specify how this distribution will evolve in time.
@@
 
-Of course, if one insists on defining a `classical system' as Curiel does, one has grounds for privileging Lagrangian mechanics.
+Of course, if one insists on defining a ``classical system'' as Curiel does, one has grounds for privileging Lagrangian mechanics.
 Nevertheless, our preferred notion of classical systems---focusing on systems that are infinitesimally reducible---has substantive advantages over Curiel's.
-Preceding sections have illustrated many of these advantages.
-Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{dependence}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a general geometrical understanding of the principle of stationary action (Section~\ref{action}).
+The general geometrical understanding of the principle of stationary action illustrates one of these advantages (Section~\ref{action}).
 Additionally, our definition accommodates models of classical massless particles by maintaining the uniqueness of solutions, even outside the hyper-regular domain.
@@
 Hence, our framework perspicuously distinguishes quantum from classical systems.
-We take these considerations as reason enough for adopting our preferred notion of a classical system.\footnote{We provide details of the quantum and statistical mechanical cases elsewhere.}
+We take these considerations to favor our preferred notion of a classical system.\footnote{We provide details of the quantum and statistical mechanical cases elsewhere.
+Moreover, we show that Hamiltonian mechanics provides the proper setting for the invariant density and a simple understanding of the gauge-dependency of conjugate momentum.}
+%Preceding sections have illustrated many of these advantages.
+Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{dependence}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a general geometrical understanding of the principle of stationary action (Section~\ref{action}).
 % % %, see \textcites{AoPPhy1}. [omit for review].
@@
 Although both frameworks adequately treat hyper-regular systems, there are compelling reasons to interpret Hamiltonian mechanics as more fundamental than Lagrangian mechanics.
-Our formalism emphasizes the fundamental significance of the invariant phase space density, along with the role of the symplectic form in counting states for each degree of freedom.
-Whereas the state variables $(q, p)$ suffer from the gauge-dependency of the conjugate momentum, the density provides an invariant description within the Hamiltonian formalism.
+Our formalism emphasizes the invariant phase space density and the role of the symplectic form in counting states for each degree of freedom.
 Finally, we have provided a geometrical interpretation of the action principle that makes the gauge-dependency of the Lagrangian manifest.
 Overall, our approach provides a better understanding of not just the relationship between Hamiltonian and Lagrangian mechanics, but of classical systems in general.
-%The importance of the density shows that a quantity can be fundamental yet coordinate-dependent.
-We have also argued that non-fundamental quantities can be physically significant despite being neither coordinate- nor gauge-invariant.
-[had to cut out since based on OG section 5.2]
-%Finally, we have provided a geometrical interpretation of the principle of stationary action that applies even outside the hyper-regular domain.
-%In fact, this explains why statistical mechanics is best done on phase space, since it provides the natural setting for an invariant density.
-[include only if space to explain this in part of section 5.]
-%We have also elucidated the role of the symplectic form in counting states for each degree of freedom.
-
-% Note that the momentum is less constrained by the position than velocity is.
-If you are given two positions at different times, then regardless of the evolutions, you can determine the velocity.
-However, from these two positions, you can be told units for the momentum but not what the momentum actually is.
-To do this, you would need to know the right relationship between  x and p.
-For instance, consider a particle that is slowing down: this could be b/c of friction, or gaining mass, or that you're in a comoving frame that is catching up.
-you can't know this just by looking at position.
-you need some additional external knowledge to distinguish these possible scenarios (e.g.
-that mass of the system is constant).
-
-%On its own, the motion of a system in physical space is not sufficient for defining the system's dynamics.
-The kinematics/motion does not suffice to determine the forces acting on the system.
-For it is essential to be able to distinguish the forces acting on an inertial system from fictitious forces arising in the description of a non-inertial system.
-Hence, unless one knows that the motion equivalence principle is satisfied, Lagrangian mechanics faces an underdetermination problem: it is unable to distinguish between certain inertial vs.
-non-inertial systems that are physically distinct.
-Distinguishing between these kinds of systems requires a relationship between conjugate momentum and kinematic momentum (or equivalently, knowledge of the energy/Hamiltonian).
-% Example of underdetermination in the context of Lagrangian mechanics: a free particle in a non-inertial frame vs.
-an accelerated particle (in an inertial frame) acted on by a dissipative force.
-These two distinct classical systems can nevertheless have the same equations of motion.
-% The Hamiltonian framework does not face a corresponding underdetermination problem because it does not allow for dissipative systems [at least when we restrict to deterministic/reversible evolN].
-Hence, there is a sense in which the Lagrangian formalism is more mathematically general than the Hamiltonian formalism.
-However, this additional generality is irrelevant when considering conservative classical systems.
-
-%Barrett also shows that the vector field formulations of Lagrangian and Hamiltonian mechanics are not categorically equivalent [1185].
-He interprets this as substantiating Curiel's claim that the ``Legendre transform does not respect the kinematical constraints of Lagrangian mechanics'', namely because it shows that ``there are Lagrangians that give rise to the same Lagrangian vector field, but do not translate---via the Legendre transformation---into Hamiltonians that give rise to the same Hamiltonian vector field'' [1186].
-What does this mean in the context of motion equivalence?
-%Barrett interprets this situation as simply showing that the Hamiltonian vector field contains more information about the Hamiltonian than the Lagrangian vector field does about the Lagrangian.
-If two Hamiltonian systems agree on the Hamiltonian vector field, then they must agree on the Hamiltonian, up to a arbitrary constant.
-Yet two Lagrangian systems can have the same Lagrangian vector field but not agree on the Lagrangian or total energy (even up to a constant).
-
-\bs
+
+\bs %begins single spacing
+
 \printbibliography[title ={References}]
@@
 
+%
+
 \end{document}

### 2021-08-05 Articles/2020-ClassicalMechanicsPhil/Submissions/PoS/classical-phil-article-PoS.tex
Update classical-phil-article-PoS.tex
Submitted version, with updated formatting/preamble.

@@
-\documentclass[12pt, english, twoside]{article} %twoside necessary for fancyheader odd vs. even commands
-
-\usepackage[letterpaper]{geometry}
-%\geometry{verbose, a4paper, tmargin=1in,bmargin=1in,lmargin=1.25in,rmargin=1.25in}
-\geometry{verbose, a4paper, tmargin=3cm,bmargin=3cm,lmargin=3cm,rmargin=3cm}
-
-\usepackage{amsmath, hyperref, enumerate, comment}
-\usepackage{amsfonts, esint, setspace}
-\usepackage{graphicx}
-\usepackage{mathrsfs} %for script font
-
-\linespread{1.3}
+%!TEX TS-program = xelatex     				
+%!TEX encoding = UTF-8 Unicode
+
+\documentclass[12pt, twoside]{article}
+\usepackage[a4paper]{geometry}
+\geometry{top=4cm, bottom=4cm, left=2.5cm, right=2.5cm} 
+%\geometry{top=4cm, bottom=4cm, left=3cm, right=4cm} %close to actual philsci margins
 
@@
 
-\usepackage[titletoc,title]{appendix}
-
-% headers
-\usepackage{fancyhdr} 
-\pagestyle{fancy}
-\fancyhead{}
-\fancyhead[OR,ER]{\large \textbf{\thepage}}
-%\fancyhead[C]{\textit{}}% odd page header and number to right top
-%\fancyhead[EC]{\textit{Your Name}}%Even page header and number at left top
-\fancyfoot[L,R,C]{}
-\renewcommand{\headrulewidth}{0pt}
-%\renewcommand{\footrulewidth}{0pt}
-\renewcommand*\footnoterule{} %removes footnote line
-
-% % redefines abstract to be capitalized, with no margins 
-\renewenvironment{abstract}
- {\small
-  \begin{center}
-\normalsize  \textnormal{ABSTRACT\\} \vspace{-0em}\vspace{0pt}
-  \end{center}
-  \list{}{%
-    \setlength{\leftmargin}{0in}%
-    \setlength{\rightmargin}{\leftmargin}%
-  }%
-  \item\relax}
- {\endlist}
-
-% section headings
-\usepackage{sectsty}
-\allsectionsfont{\large\raggedright\centering}
+\usepackage{footmisc} % all footnote commands are at end of preamble
 
@@
 
-% % footnote modification: must be between \makeatletter and \makeatother
-
-\renewcommand{\@makefntext}[1]{%
-  \setlength{\parindent}{0pt}%
-  \begin{list}{}{\setlength{\labelwidth}{6mm}% indents footnote text 
-    \setlength{\leftmargin}{\labelwidth}%
-    \setlength{\labelsep}{5pt}% moves footnote mark to the left
-     \setlength{\itemsep}{0pt}%
-      \setlength{\parsep}{0pt}%
-      \setlength{\topsep}{-3pt}% removes separation between footnotes
-    \footnotesize}%
-  \item[\@textsuperscript{\@thefnmark}\hfil ]#1% @makefnmark
-  \end{list}%
-}
-%      \setlength{\footnotesep}{0pt}%
-
-\makeatother
-\renewcommand{\cftdot}{} % removes dots from table of contents
+\usepackage{titlesec}
+\iffalse
+\titleformat{\section}[runin]% runin puts it in the same paragraph
+        {\normalfont\bfseries}% formatting commands to apply to the whole heading
+        {\thesection.}% the label and number
+        {0.5em}% space between label/number and section title
+        {}% formatting commands applied just to section title
+        [.]% punctuation or other commands following section title
+
+\titlespacing*{\section}{0pt}{5mm}{2.5mm}%spacing from left margin, spacing above header, spacing on right of header
+\fi
+
+% headers
+\usepackage{fancyhdr} 
+\pagestyle{fancy}
+\fancyhead{}
+\fancyhead[R]{\thepage} %page number placement
+\fancyhead[C]{\textsc{HAMILTONIAN PRIVILEGE}}% odd page header 
+%\fancyhead[EC]{\textsc{AUTHOR NAMES}}%Even page header 
+\fancyfoot[L,R,C]{}
+\renewcommand{\headrulewidth}{0pt}
+%\renewcommand{\footrulewidth}{0pt}
+%\renewcommand*\footnoterule{} % removes footnote line
+
+\fancypagestyle{plain} % for first page 
+{
+\fancyhead[LCR]{} %clears header 
+\fancyfoot[C]{\thepage}
+   \renewcommand{\headrulewidth}{0pt}
+}
+
+\usepackage{amsmath, hyperref, enumerate, comment}
+\usepackage{amsfonts, esint, setspace}
+\usepackage{graphicx}
+\usepackage{mathrsfs} %for script font
+
+%\doublespacing
+\linespread{1.5}
 
@@
 \usepackage{mathptmx}
-%\usepackage{mathpazo}
+
+% % Generates bib file for project: %place before biblatex commands and calls
+\usepackage{filecontents}
+\begin{filecontents}{references.bib}
+
+% Encoding: UTF-8
+
+@book{Arnold,
+	author = {Arnold, Vladimir I.},
+	title = {Mathematical Methods of Classical Mechanics},
+	year = {1989},
+	publisher = {Springer},
+	place = {New York},
+	edition = {2},
+	translator = {K. Vogtmann and A. Weinstein}
+}
+
+@book{Abraham,
+	author = {Abraham, Ralph and Marsden, Jerrold E.},
+	title = {Foundations of Mechanics},
+	year = {1978},
+	publisher = {Benjamin--Cummings Publishing Company},
+	place = {London},
+	edition = {2},
+}
+
+@Unpublished{Butterfield2004,
+  author  = {Butterfield, Jeremy N.},
+  Title   = {Between Laws and Models: {S}ome Philosophical Morals of {L}agrangian Mechanics},
+  Year    = {2004},
+  Address = {https://arxiv.org/abs/physics/0409030},
+}
+
+@InCollection{Butterfield2004a,
+  Title     = {On {H}amilton--{J}acobi Theory as a Classical Root of Quantum Theory},
+  Publisher = {Springer},
+  Year      = {2004},
+  author  = {Butterfield, Jeremy N.},
+  Editor    = {Elitzur, A. C. and Dolev, S. and Kolenda, N.},
+  Address   = {New York},
+  Booktitle = {Quo Vadis Quantum Mechanics? {P}ossible Developments in Quantum Theory in the 21st Century},
+}
+
+@Article{Butterfieldpoints,
+  author  = {Butterfield, Jeremy N.},
+  journal = {British Journal for Philosophy of Science},
+  title   = {Against Pointillisme about Mechanics},
+  year    = {2006},
+  pages   = {709--753},
+  volume  = {57},
+}
+
+@InCollection{Butterfield2006,
+  Pages     = {43--99},
+  Title     = {On Symmetry and Conserved Quantities in Classical Mechanics},
+  Publisher = {Springer},
+  Year      = {2006},
+  Author    = {Butterfield, Jeremy N.},
+  Editor    = {Demopoulos, William and Pitowsky, Itamar},
+  Address   = {Dordrecht},
+  Booktitle = {Physical Theory and its Interpretation: {E}ssays in Honor of {J}effrey {B}ub},
+}
+
+@book{Cline,
+	author = {Cline, Douglas},
+	title = {Variational Principles in Classical Mechanics},
+	year = {2021},
+	publisher = {LibreTexts},
+}
+
+@Book{Field,
+  Title     = {Science without Numbers: {A} Defense of Nominalism},
+  Address   = {Oxford},
+  Publisher = {Oxford University Press},
+  Year      = {2016 [1980]},
+edition = {2},
+  Author    = {Field, Hartry H.},
+}
+
+@Book{Field1980,
+  Title     = {Science without Numbers: {A} Defence of Nominalism},
+  Publisher = {Princeton University Press},
+  Year      = {1980},
+  Author    = {Field, Hartry H.},
+  Address   = {Princeton, N.J.},
+  Note      = {80007560 Hartry H. Field. Ill. ; 23 Cm. Includes Bibliographical References and Index.},
+  Keywords  = {Science Philosophy. Mathematics Philosophy. Nominalism.},
+  Pagetotal     = {xiii, 130 P.},
+}
+
+@Book{Healey,
+  Title     = {Gauging What's Real: The Conceptual Foundations of Contemporary Gauge Theories},
+  Publisher = {Oxford University Press},
+  Year      = {2007},
+  Author    = {Healey, Richard},
+  Address   = {Oxford},
+}
+
+@Book{Ruetsche,
+  Title     = {Interpreting Quantum Theories},
+  Address   = {Oxford},
+  Publisher = {Oxford University Press},
+  Year      = {2011},
+  Author    = {Ruetsche, Laura},
+}
+
+@Article{Smith,
+  Author    = {Smith, Sheldon R},
+  Title     = {Symmetries and the Explanation of Conservation Laws in the Light of the Inverse Problem in {L}agrangian Mechanics},
+  Journal   = {Studies in History and Philosophy of Modern Physics},
+  Year      = {2008},
+  Volume    = {39},
+  Number    = {2},
+  Pages     = {325--345},
+  Publisher = {Elsevier},
+}
+
+@ARTICLE {North,
+	author  = "Jill North",
+	title   = "The ``Structure'' of Physics: A Case Study",
+	journal = "The Journal of Philosophy",
+	year    = "2009",
+	volume  = "106",
+	number  = "2",
+	pages   = "57--88",
+	url     = "http://philsci-archive.pitt.edu/4961/"
+}
+
+@ARTICLE {Curiel,
+	author  = "Erik Curiel",
+	title   = "Classical Mechanics Is Lagrangian; It Is Not Hamiltonian",
+	journal = "The British Journal for the Philosophy of Science",
+	year    = "2014",
+	volume  = "65",
+	number  = "2",
+	pages   = "269-–321",
+	url     = "https://doi.org/10.1093/bjps/axs034"
+}
+
+@ARTICLE {Barrett1,
+	author  = "Thomas William Barrett",
+	title   = "On the Structure of Classical Mechanics",
+	journal = "The British Journal for the Philosophy of Science",
+	year    = "2014",
+	volume  = "66",
+	number  = "4",
+	pages   = "801-–828",
+	url     = "https://doi.org/10.1093/bjps/axu005"
+}
+
+@ARTICLE {Barrett2,
+	author  = "Thomas William Barrett",
+	title   = "Equivalent and Inequivalent Formulations of Classical Mechanics",
+	journal = "The British Journal for the Philosophy of Science",
+	year    = "2018",
+	url     = "https://doi.org/10.1093/bjps/axy017"
+}
+
+@ARTICLE {AllHamFreeParticle,
+	author  = "E. Galindo-Linares and E. Navarro-Morale and G. Silva-Ortigoza and R. Suárez-Xique and M. Marciano-Melchor and R. Silva-Ortigoza and E. Román-Hernández",
+	title   = "Any Hamiltonian System Is Locally Equivalent to a Free Particle",
+	journal = "World Journal of Mechanics",
+	year    = "2012",
+	volume  = "2",
+	number  = "5",
+	pages   = "246--252",
+	url     = "https://doi.org/10.4236/wjm.2012.25030"
+}
+
+@ARTICLE {AllSystemsAreHam,
+	author  = "C. A. Lucey and E. T. Newman",
+	title   = "On the construction of Hamiltonians",
+	journal = "Journal of Mathematical Physics",
+	year    = "1988",
+	volume  = "29",
+	number  = "11",
+	pages   = "2430--2433",
+	url     = "https://doi.org/10.1063/1.528129"
+}
+
+@article{Shannon,
+	added-at = {2018-06-03T07:16:08.000+0200},
+	author = {Shannon, Claude Elwood},
+	biburl = {https://www.bibsonomy.org/bibtex/2991e1d12fe529405c363bd2aad283d7b/achakraborty},
+	doi = {10.1002/j.1538-7305.1948.tb01338.x},
+	interhash = {754130207906fcec16a53d330eeff348},
+	intrahash = {991e1d12fe529405c363bd2aad283d7b},
+	journal = {The Bell System Technical Journal},
+	keywords = {ieee information-theory paper},
+	month = {7},
+	number = 3,
+	pages = {379--423},
+	publisher = {Nokia Bell Labs},
+	timestamp = {2018-06-03T07:17:55.000+0200},
+	title = {A Mathematical Theory of Communication},
+	url = {https://ieeexplore.ieee.org/document/6773024/},
+	volume = 27,
+	year = 1948
+}
+
+@article{Jaynes,
+	added-at = {2017-05-10T05:56:13.000+0200},
+	author = {Jaynes, Edwin T},
+	biburl = {https://www.bibsonomy.org/bibtex/2521dc36203688bc4c8b065e18dea3a7c/salotz},
+	interhash = {cbe997458351b8d62a904b58205078ea},
+	intrahash = {521dc36203688bc4c8b065e18dea3a7c},
+	journal = {Physical review},
+	keywords = {information-theory jaynes statistical-mechanics},
+	number = 4,
+	pages = 620,
+	publisher = {APS},
+	timestamp = {2017-05-10T05:56:13.000+0200},
+	title = {Information theory and statistical mechanics},
+	volume = 106,
+	year = 1957
+}
+
+@article{Holten,
+	title = {Covariant Hamiltonian dynamics},
+	author = {{van Holten}, J. W.},
+	journal = {Phys. Rev. D},
+	volume = {75},
+	issue = {2},
+	pages = {025027},
+	numpages = {9},
+	year = {2007},
+	month = {Jan},
+	publisher = {American Physical Society},
+	doi = {10.1103/PhysRevD.75.025027},
+	url = {https://link.aps.org/doi/10.1103/PhysRevD.75.025027}
+}
+
+@article{Struckmeier,
+	title = {Hamiltonian Dynamics on the Symplectic Extended Phase Space for Autonomous and Non-autonomous Systems},
+	author = {Struckmeier, J.},
+	journal = {Journal of Physics A: {M}athematical and General},
+	volume = {38},
+	number = {6},
+	pages = {1257--1278},
+	year = {2005},
+	doi = {10.1088/0305-4470/38/6/006},
+	url = {https://iopscience.iop.org/article/10.1088/0305-4470/38/6/006},
+}
+
+@article{Coffey,
+	author = {Coffey, Kevin},
+	year = {2014},
+	month = {11},
+	pages = {821-844},
+	title = {Theoretical Equivalence as Interpretative Equivalence},
+	volume = {65},
+	journal = {The British Journal for the Philosophy of Science},
+	doi = {10.1093/bjps/axt034}
+}
+
+@book{Maudlin,
+	author = {Tim Maudlin},
+	title = {New Foundations for Physical Geometry: The Theory of Linear 	Structures.},
+	year = {2014},
+	publisher = {Oxford University Press},
+	place = {New York}
+}
+
+@article{Jarzynski,
+	title = {Stochastic and Macroscopic Thermodynamics of Strongly 	Coupled Systems},
+	author = {Jarzynski, Christopher},
+	journal = {Phys. Rev. X},
+	volume = {7},
+	issue = {1},
+	pages = {011008},
+	numpages = {20},
+	year = {2017},
+	month = {Jan},
+	publisher = {American Physical Society},
+	doi = {10.1103/PhysRevX.7.011008},
+	url = {https://link.aps.org/doi/10.1103/PhysRevX.7.011008}
+}
+
+@article{Maes,
+	title={Time-reversal and entropy},
+	author={Maes, Christian and Neto{\v{c}}n{\`y}, Karel},
+	journal={Journal of statistical physics},
+	volume={110},
+	number={1-2},
+	pages={269--310},
+	year={2003},
+	publisher={Springer}
+}
+
+@article{Horowitz,
+	title={Thermodynamics of information},
+	author={Parrondo, Juan MR and Horowitz, Jordan M and 	Sagawa, Takahiro},
+	journal={Nature physics},
+	volume={11},
+	number={2},
+	pages={131},
+	year={2015},
+	publisher={Nature Publishing Group}
+}
+@Article{Teh,
+  author  = {Teh, Nicholas and Tsementzis, Dimitris},
+  journal = {Studies in History and Philosophy of Modern Physics},
+  title   = {Theoretical Equivalence in Classical Mechanics and its Relationship to Duality},
+  year    = {2017},
+  pages   = {44--54},
+  volume  = {59},
+}
+
+@Unpublished{Swanson,
+	author = {Swanson, Noel and Halvorson, Hans},
+	year = {2012},
+	month = {},
+	pages = {1-10},
+	title = {On North's `The Structure of Physics'},
+	volume = {},
+	journal = {},
+	doi = {},
+	address = {http://philsci-archive.pitt.edu/9314/}, 
+}
+
+@Article{Lewis1983,
+  Author    = {Lewis, David},
+  Title     = {New Work for a Theory of Universals},
+  Journal   = {Australasian Journal of Philosophy},
+  Year      = {1983},
+  Volume    = {61},
+  Number    = {4},
+  Pages     = {343--377},
+  Publisher = {Taylor \& Francis Group},
+}
+
+@Book{Earman,
+  author    = {Earman, John},
+  publisher = {Reidel},
+  title     = {A Primer on Determinism},
+  address   = {Dordrecht},
+  year      = {1986},
+}
+
+@Unpublished{Baez,
+  author = {Baez, John C.},
+  title  = {Struggles with the Continuum},
+  year   = {unpublished},
+  url    = {https://arxiv.org/abs/1609.01421},
+  address = {https://arxiv.org/abs/1609.01421}, 
+}
+
+@Article{Fletcher,
+  author    = {Fletcher, Samuel Craig},
+  journal   = {European Journal for Philosophy of Science},
+  title     = {What Counts as a Newtonian System? the View from Norton’s Dome},
+  year      = {2012},
+  month     = Oct,
+  number    = {3},
+  pages     = {275},
+  volume    = {2},
+  date      = {2012-10-01},
+  doi       = {10.1007/s13194-011-0040-8},
+  publisher = {Springer},
+  url       = {http://dx.doi.org/10.1007/s13194-011-0040-8},
+}
+
+@Article{Norton,
+  author  = {Norton, John},
+  journal = {Philosophy of Science},
+  title   = {The {D}ome: {A}n Unexpectedly Simple Failure of Determinism},
+  year    = {2008},
+  number  = {5},
+  pages   = {786--798},
+  volume  = {75},
+}
+
+@InCollection{Dorr_Hawthorne,
+  author    = {Dorr, Cian and Hawthorne, John},
+  booktitle = {Oxford Studies in Metaphysics},
+  publisher = {Oxford University Press},
+  title     = {Naturalness},
+  year      = {2014},
+  editor    = {Bennett, Karen and Zimmerman, Dean W.},
+  pages     = {3--77},
+  volume    = {8},
+}
+
+@InCollection{Wilson,
+  author    = {Wilson, Mark},
+  booktitle = {The Oxford Handbook of Philosophy of Physics},
+  publisher = {Oxford University Press},
+  title     = {What is `Classical Mechanics' Anyways?},
+  year      = {2013},
+  address   = {Oxford},
+  chapter   = {2},
+  editor    = {Batterman, Robert},
+  pages     = {43--106},
+}
+
+@InCollection{Wallace,
+  author    = {Wallace, David},
+  booktitle = {Symmetries in Physics: {P}hilosophical Reflections},
+  publisher = {Cambridge University Press},
+  title     = {Time-Dependent Symmetries: {T}he Link between Gauge Symmetries and Indeterminism},
+  year      = {2003},
+  address   = {Cambridge},
+  editor    = {Brading, Katherine and Castellani, Elena},
+  pages     = {163--173},
+}
+
+@book{Yablo,
+	author = {Yablo, Stephen},
+	title = {Aboutness},
+	year = {2014},
+	publisher = {Princeton University Press},
+	address = {Princeton},
+}
+
+@InCollection{Belot,
+  author    = {Belot, Gordon},
+  booktitle = {The Oxford Handbook of Philosophy of Physics},
+  publisher = {Oxford University Press},
+  title     = {Symmetry and Equivalence},
+  year      = {2013},
+  address   = {Oxford},
+  chapter   = {9},
+  editor    = {Batterman, Robert},
+  pages     = {318--339},
+}
+
+@Article{Chen2018,
+  author  = {Eddy Keming Chen},
+  title   = {The Intrinsic Structure of Quantum Mechanics},
+  year    = {unpublished},
+url = {https://arxiv.org/abs/1810.06551},
+}
+
+\end{filecontents}
+
+% % % Biblatex commands for Citations and References:
+\usepackage[backend=bibtex, authordate, url=false, doi=false, eprint=false]{biblatex-chicago}
+%\addbibresource{bibliography.bib}
+\addbibresource{references.bib} %call the bib file generated above
+
+% clears un-needed fields from reference items
+\AtEveryBibitem{%
+  \clearfield{note}%
+  \clearfield{series}%
+  \clearfield{pagetotal}%
+   \clearfield{chapter}%
+  \clearfield{number}%for convenience/consistency
+}
+
+\DeclareFieldFormat{postnote}{\mkcomprange{#1}} % compresses page range in citations, so that unchanged digits are not repeated. See p. 259, section 4.6.4 of biblatex manual.
+
+\DeclareFieldFormat{pages}{\mkcomprange{#1}} % compresses page range in bibliography, so that unchanged digits are not repeated.
+
+\DefineBibliographyStrings{english}% for your language
+{ byeditor = {ed.}, } %changes `Edited by' to `ed.' See Stanley easier guide, p. 35
+
+%\renewcommand*{\bibfont}{\footnotesize} %changes fontsize of bibitems
+
+\newcommand\bs{\begin{singlespace}} 			
+\newcommand\es{\end{singlespace}}
+
+%Physics notation 
+\newcommand{\diff}{\mathop{}\!\mathrm{d}} %defines exterior derivative 
+\renewcommand{\div}[1]{\gv{\nabla} \cdot #1} % for divergence
+\newcommand{\curl}[1]{\gv{\nabla} \times #1} % for curl
+\renewcommand{\vector}[1]{\ensuremath{\vec{#1}}} % redefines vector to vec
+\newcommand{\integral}{\int}
 
@@
 \hypersetup{
-	pdftitle={Your Paper Title},
-	pdfauthor={Your Name},
+	pdftitle={Hamiltonian Privilege},
+	pdfauthor={},
 	pdfsubject={},
@@
 	colorlinks=true,	% false: boxed links; true: colored links
-	linkcolor=black,	% color of internal links
+	linkcolor=blue,	% color of internal links
 	citecolor=black,	% color of links to bibliography
 	filecolor=black,	% color of file links
-	urlcolor=blue		% color of external links
-}
-
-%\usepackage[backend=biber, style=authoryear, url=false, doi=false, eprint=false]{biblatex}
-
-% % % Biblatex commands for Citations and References:
-\usepackage[backend=biber, style=ext-authoryear-comp,  maxcitenames=2, dashed=false, url=false, doi=false, eprint=true, giveninits=true, innamebeforetitle=true, maxbibnames=6]{biblatex}
-
-\addbibresource{bibliography.bib}
-
-%\addbibresource{templateReferences.bib} %%generated by reference items above
-%maxcitenames=2 replaces all instances of more than 2 authors in citations with `et al.', but lists all authors in references
-%labeltitleyear=true; sortcites=true
-%innamebeforetitle=true places editors before the book titles
-%style=authoryear-comp adds in some weird ``In:'' stuff in reference list that must be changed below.
-
-%\DeclareFieldFormat{labeldate}{\mkbibbrackets{#1}} %puts brackets around years in citations
-
-\DeclareFieldFormat{postnote}{\mkpageprefix[pagination][\mkcomprange]{#1}} % compresses page range in citations, so that unchanged digits are not repeated.
-See p.
-259, section 4.6.4 of biblatex manual.
-%also adds pagination prefix p.
-
-% clears un-needed fields from reference items
-\AtEveryBibitem{%
- \clearfield{note}%
-  \clearfield{series}%
- \clearfield{number}%
- \clearfield{pagetotal}%
- \clearfield{chapter}%
-  }
-  %\clearfield{pages}%}
-
-  \renewcommand{\newunitpunct}{\addcomma\space} %makes a comma the default separator between items; overridden in some cases by commands below
-
-\renewcommand{\intitlepunct}{\addspace\nopunct} %removes punctuation after ``in"
-
-\DeclareDelimFormat[bib,biblist]{nametitledelim}{\addcolon\space} %puts colon after year
-
-%\DeclareFieldFormat{biblabeldate}{\mkbibbrackets{#1}} %puts brackets around years in references
-
-\DeclareFieldFormat{editortype}{\mkbibparens{\textit{#1}}} %puts ed in parentheses
-\DeclareDelimFormat{editortypedelim}{\addspace}%prevents punctuaction clash after (ed)
-\DeclareDelimFormat[bib,biblist]{innametitledelim}{\addcomma\space} % puts comma after (ed)
-
-\DeclareFieldFormat{eprint}{\printtext{available at}\addspace \textless#1\textgreater} % Calls url from Eprint field. See biblatex manual, section 4.11.2 Electronic Publishing Information, page 310-11
-
-\DeclareFieldFormat[article,periodical]{volume}{\textbf{#1}} %puts volume in bold; note that adding \bibstring{jourvol}~ before #1} prints ``vol.". from biblatex-ext manual 
-%\DeclareFieldFormat[article,periodical]{number}{#1}%\bibstring{number}~ before #1} prints ``vol."
-\renewcommand*{\jourvoldelim}{\addcomma\space} %combined with above command, puts comma after article title and before journal volume. 
-%\renewcommand*{\volnumdelim}{\addcomma\space}
-
-\DeclareFieldFormat[inbook, incollection, book]{volume}{\bibcpstring{volume}~#1} %capitalizes `Vol' after book titles.
-
-\DeclareFieldFormat{pages}{\mkpageprefix[pagination][\mkcomprange]{#1}} % compresses page range in bibliography, so that unchanged digits are not repeated.
-
-% % following command suppresses ``in" for articles and inproceedings only
-\renewbibmacro{in:}{%
-\ifboolexpr{%
-test {\ifentrytype{article}}%
-or
-test {\ifentrytype{inproceedings}}%
-}
-{}%<-nothing
-{\printtext{\bibstring{in}\intitlepunct}}%<-normal ’in’ with punctuation
-}
-
-\newcommand{\diff}{\mathop{}\!\mathrm{d}} %defines exterior derivative 
-\renewcommand{\div}[1]{\gv{\nabla} \cdot #1} % for divergence
-\newcommand{\curl}[1]{\gv{\nabla} \times #1} % for curl
-\renewcommand{\vector}[1]{\ensuremath{\vec{#1}}} % redefines vector to vec
-\newcommand{\integral}{\int}
-
-\newcommand\bs{\begin{singlespace}} 			
-\newcommand\es{\end{singlespace}}
+	urlcolor=black		% color of external links
+}
+
+\newif\iffootpunct %defines \footpunctfalse  and \footpuncttrue commands
+
+% % Essential to modify footnote ONLY AFTER BIBLATEX COMMANDS, since otherwise ruins citation styles
+
+% % footnote modification: must be between \makeatletter and \makeatother
+ \makeatletter
+
+ \def\blfootnote{\gdef\@thefnmark{} \@footnotetext} %footnote without number marker
+
+ \renewcommand{\@makefntext}[1]{%
+  \setlength{\parindent}{0pt}%
+  \begin{list}{}{\setlength{\labelwidth}{0mm}% indents footnote text, Space from number to border
+    \setlength{\leftmargin}{\labelwidth}%  
+    \setlength{\labelsep}{0pt}% Space from number to text, moves footnote mark to the left
+     \setlength{\itemsep}{0pt}%
+      \setlength{\parsep}{0pt}%
+      \setlength{\topsep}{0pt}% removes separation between footnotes
+      }%\doublespacing \normalsize% double spaces and changes font size of footnotes; place before }
+  \item[\@thefnmark \iffootpunct. \fi \hfil ]#1% @makefnmark, with period iff footpuncttrue
+  \end{list}%
+}
+
+%      \setlength{\footnotesep}{0.83125\baselineskip}% separation between footnotes
+
+\makeatother
+
+%\renewcommand{\footnotelayout}{\normalsize}
 
@@
 %\author{\Large \textsuperscript{*\textdagger}}
-\date{\vspace{-7.9ex} \today}
+\date{\vspace{-7.9ex} July 30, 2021}
 \maketitle
@@
 \es
+
+\footpunctfalse% turns off period after footnote marker for blindfootnote
+
+%\blfootnote{*To contact the author, please write to: }
+%\blfootnote{\textdagger For comments on earlier drafts, we thank }
+
+\footpuncttrue% turns back on the period after the footnotemarker
+
+\raggedright\setlength{\parindent}{1em}
 
@@
 Hence, our framework perspicuously distinguishes quantum from classical systems.
-We take these considerations to favor our preferred notion of a classical system.\footnote{We provide details of the quantum and statistical mechanical cases elsewhere.
+We take these considerations to favor our preferred notion of a classical system.\footnote{Space precludes a detailed discussion of the quantum and statistical mechanical cases, which we hope to discuss elsewhere.
+Additionally, one can show that Hamiltonian mechanics provides the proper setting for the invariant density and a simple understanding of the gauge-dependency of conjugate momentum.}
+
+%\footnote{We provide details of the quantum and statistical mechanical cases elsewhere.
 Moreover, we show that Hamiltonian mechanics provides the proper setting for the invariant density and a simple understanding of the gauge-dependency of conjugate momentum.}
@@
 
+\clearpage
+
 \bs %begins single spacing
 
-\printbibliography[title ={References}]
+\setlength\bibitemsep{2pt} %controls spacing between bib entries. default is itemsep
+
+\defbibheading{myheading}[REFERENCES]{%
+\centering \normalfont\textsc{{#1}}%
+%\centering \footnotesize\textsc{{#1}}%
+%\addcontentsline{toc}{subsection}{#1}% if you want this in table of contents; would need to use a \section* command above as well 
+%\markboth{#1}{#1}
+}
+
+\printbibliography[heading=myheading]%see Stanley easier read, pages 63-64.
+
+%\printbibliography[title ={\centering \normalfont \textsc{References}}]
 \es

### 2021-08-06 Overviews/Differentiability/Differentiability.tex
Differentiability: first notes

\documentclass[10pt, onecolumn, longbibliography, nofootinbib]{revtex4-2}

\usepackage{assumptionsofphysics}

\usepackage{tikz}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\frenchspacing

% Usual (decimal) numbering
\renewcommand{\thesection}{\arabic{section}}
\renewcommand{\thesubsection}{\thesection.\arabic{subsection}}
\renewcommand{\thesubsubsection}{\thesubsection.\arabic{subsubsection}}

% Fix references
\makeatletter
\renewcommand{\p@subsection}{}
\renewcommand{\p@subsubsection}{}
\makeatother

\begin{document}

\title{Physical motivation for differentiability}
%\author{Gabriele Carcassi}
%\email{carcassi@umich.edu}
%\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
%\author{Christine Aidala}
%\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\date{\today}

\begin{abstract}
	This paper presents ideas and progress towards a more direct physical justification for differentiability. The general idea that the differentiable structure on a manifold exists to allow us to express a distribution as a density over the coordinates.
\end{abstract}

\maketitle

\section{Introduction}

\emph{This paper presents the goal and the current status of our research. It is intended to gather early feedback, including pointers to other relevant work. Feedback is welcome and encouraged.}

As part of our larger project, Assumptions of Physics, we want to rederive the mathematical structures we use in physics from precise physical starting points. In previous works we have shown the following:
\begin{description}
	\item[Experimentally distinguishable cases] $\to$ topological space. Each open set corresponds to an experimentally testable hypothesis.
	\item[Cases identifiable by continuous quantities] $\to$ manifold. Coordinates represent measurable quantities.
\end{description}
Before we establish what Riemannian and symplectic manifolds correspond to, we need to establish what differentiability corresponds to.

The general idea is that differentiability allows us to express a distribution as a continuous function $\rho(x^i)$ of the coordinates. That is, among all the possible coordinates of the manifold we restrict ourselves to those that allow us to express $\rho$ in a nice way.

Conceptually, we will want to start from statements that are directly testable. As we do not measure densities directly, but finite values in finite regions, the density distribution $\rho$ cannot be our starting point. Given a set $X$ of possible cases, our verifiable statements are of the form ``in region $U$ the value of the quantity is within the range $V$'', where $U$ is an open set of $X$ and $V$ is an open set of $\mathbb{R}$. We need to identify the exact conditions for which we will recover differentiability in the case that $X$ is a manifold.

\section{Mathematical problem}

\begin{defn}
	Let $X$ be a manifold, $\textsf{T}_X$ its topology and $\Sigma_X$ its Borel algebra. A \textbf{distributed quantity} $\mu$ is a measure $\mu : \Sigma_X \to \mathbb{R}$ such that $\mu(U) = \mu(\interior(U))$. 
\end{defn}

\begin{defn}
	Given a manifold $X$ and a distributed quantity $\mu$, we say a coordinate $q : U \to \mathbb{R}$ is \textbf{smooth with respect to $\mu$} if the Radon-Nikodym derivative is continuous.
\end{defn}

\begin{desid}
	The set of all smooth coordinates with respect to a distributed quantity $\mu$ identifies a differentiable structure for $X$.
\end{desid}

%\section{Examples}

%\section{Insightful failures}

\bibliography{bibliography}

\end{document}

### 2021-08-10 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
ClassicalFail: first draft reviewed with Christine

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amsfonts}

\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\begin{document}

\title{Why classical mechanics fails}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
We present the idea that, contrary to the established view, classical mechanics fails for conceptual reasons, not merely because it does not match experimental evidence. The notion of a classical system stems from two underlying assumptions, perfect isolation and infinitesimal divisibility, that are in contradiction: at some point we will reach a scale for which the parts can no longer be assumed to be isolated. This result defines more cleanly the realm of applicability of classical mechanics, shows what problems quantum mechanics addresses and gives hints as to what other problems will need to be solved by a more general theory.
\end{abstract}

\maketitle

\section{Introduction}

The common view is that classical mechanics is a perfectly consistent theory that just does not happen to be true. It fails simply because experiments do not agree with it. This view is so established, so ingrained, that it is seldom discussed explicitly, and we believe it to be part of the reason why quantum mechanics still feels, after almost a century, so mysterious: how are we going to conceptually appreciate a theory if we do not even know what conceptual problems it is trying to address?

Our argument can be schematically summarized as follows:
\begin{enumerate}
	\item To define a system, we need to define a clear boundary between the system and everything else
	\item The interaction across the boundary participates in the definition of the system
	\item Perfect characterization of a system requires perfect isolation
	\item As we get to smaller and smaller parts, perfect isolation becomes untenable
\end{enumerate}
In a nutshell, classical mechanics fails because the definition of a system is too simplistic. It fails to recognize the role the system/environment interaction plays in the characterization of the system and its state space.

We will analyze these ideas first conceptually and how they translate into inconsistencies between classical mechanics on one side and thermodynamics/statistical mechanics on the other. We will then show how quantum mechanics addresses some of these issues and how it leaves some open, to be solved in future theories.

\section{System/environment boundary}

When defining a physical system, we need to define, implicitly or explicitly, its boundary. Whether we are studying a solid object, an open container of gas, a chemical compound or a single particle, we need, at the very least, to be clear as to what is in the system and what is out. This much should be evident. What may be less evident is that the nature of the interaction at the boundary between the system and the environment is also important, not just in thermodynamics and statistical mechanics, but more broadly to all physical systems. Let us start with the following simple example.

Suppose we want to study the motion of a cannon ball. As the cannon ball moves, air molecules will scatter off its surface. Yet, the effect produced will be negligible. We can therefore describe the state of the cannon ball with a single, infinitely precise, pair of position and momentum. Now suppose we want to study the motion of a speck of dust. This time the effect of the air molecules will not be negligible. We will therefore describe the state of the speck of dust with a probability distribution over position and momentum. Now suppose we want to study the motion of the cannon ball on the surface of the sun. Plasma will scatter off the ball catastrophically, destroying the ball.

The example illustrates that the system, the interaction at the boundary and the state space for the system are not three independent choices. One may constrain the others. A cannonball at 5000 Kelvin does not exist, or at least not for long. Conceptually, we can think of a property as being intrinsic to the system only if the interaction with the environment leaves it, at least to a good approximation, undisturbed. If it is not left undisturbed, then it is no longer a property of the system itself: it is a shared property between the system and the environment because the nature of both comes into play in its definition.

While we are used to think this way for thermodynamic quantities, like temperature or pressure, or statistical quantities, like average energy or mean free path, we typically think of mechanical properties, like position and velocity, differently. These, we think, are always intrinsic to the system. However, we have no basis for that distinction. What happens is that, in the vast number of cases, we can sufficiently isolate the system so that the effect of the environment on the motion of the body is negligible. We can take a speck of dust, put it in a vacuum, and the position is now much more stably definable. Therefore, when studying the speck in a fluid, the motion is modeled as fluctuations of an infinitely precise quantity because, at least conceptually, we could instantly remove the air and determine the position of the speck in isolation. But, realistically, we can only go so far. Perfect isolation is impossible.

Apart from practical concerns, any wall we will use as shielding will necessarily have finite temperature and will emit thermal radiation; moreover, gravitational interactions cannot be shielded in the first place. Even theoretically, complete isolation is not possible. We cannot, even conceptually, instantly remove all the background interactions. If that is the case, we have to concede that even properties like position and momentum are not intrinsic to the system: they are the characterization of the system together with the environment (e.g. the gravitational field).

One may still argue that these are still circumstantial problems that derive from the incompleteness of our current understanding. We could always imagine that we will be able, in the future, to achieve perfect isolation. But let us explore that possibility in detail. Perfect isolation would mean that no signal, electromagnetic, gravitational or otherwise, would be able to interact with the object. Then, how are we supposed to measure anything about the system? It is experimentally inaccessible. Maybe we are already surrounded by perfectly isolated systems: we have no way of knowing.

This issue uncovers another couple of hidden assumptions. First of all, the observer is always part of the environment as it is not part of the system. The second is that a measurement has to cross the boundary. The act of measurement may interfere with the previous nature of the system/environment interaction, which as we saw participates in defining the properties of the system itself. Therefore the act of measurement may change the nature of the very properties that we are going to measure.

Note that these are definitional issues, not mechanistic issues. It is not about what the system is doing, what the rules are that the most fundamental objects follow. It is about how we define the system in the first place: how we carve up the world into parts that we can study independently. If we take a physical system simply as something we want to study under a set of conditions, and the state to be the experimentally accessible information that is well defined under such conditions, these issues will always crop up. They need to be addressed, even if just by simplifying assumptions, when the system is being defined.

\section{Classical failure}

We are now in a position to understand the failure of classical theories. As we saw before, we can assign intrinsic properties to systems that can be, at least in principle, isolated perfectly. For large scale objects, this is not a problem: small perturbations will have small effects which can, at least at a certain scale, be neglected. But this is an idealization. The infinite precision of a real number is valid only insofar that the idealization holds.

Assuming perfect isolation by itself is not an issue. However, classical mechanics also assumes, implicitly, that this perfect isolation holds for objects of all sizes. With continuous materials, we can take smaller and smaller parts. Classical fields allow for wave packets of ever decreasing energy. The theory implicitly assumes all these objects can be prepared, manipulated and detected independently enough that their state can be perfectly defined. This is the contradiction inherent to classical mechanics: perfect isolation at all scales.

The disconnect is that many take the classical state to be the ``real'' state of the system. But this is more of a philosophical construct. What physics should be concerned with is only the part of reality that is experimentally accessible under suitable conditions (e.g. repeatability, independence from other systems, ...). This does not mean that objects we cannot study experimentally do not exist or are not worth studying. It just means that our physical theories should not make any claims about them as they are not the subject of physics.

Since this is a point of contention, and we do not believe further discussion along these lines will change anyone's perspective, let us switch gears. We are going to argue that a perfect description of the type provided by classical mechanics is inconsistent with thermodynamics and statistical mechanics.

We start by asking: what is the entropy of a single microstate? This is complicated by the fact that we have (at least) two definitions of entropies, and people argue about which is the correct one. Therefore let us use both definitions and see what happens.

Recall that a microstate for a classical system of $n$ particles is a point in a $\mathbb{R}^{6n}$ space, that represents all their three-dimensional position and momenta. The fundamental postulate of statistical mechanics states that the entropy is $S = k \log W$, where $W$ measures the  microstates. At first glance, since we have a single microstate, we are inclined to say $W = 1$. This would be true for a discrete state space. However, for a continuum state space this is the wrong answer.\footnote{This is an example where the difference between discrete and continuous spaces plays a fundamentally important role in physics. Unfortunately, too many physicists confuse discrete with countable.} We need to use the phase space volume. Since we have a single point, its volume is zero, therefore we have $S = k \log 0 = -\infty$: the entropy of a single microstate is minus infinity.

Let us now confirm the result with the Gibbs/Shannon entropy $S = -\int \rho \log \rho \, dq^{3n} dp^{3n}$. A single microstate would be described by a delta distribution centered on that microstate: $\delta(q^i-q^i_0, p^i-p^i_0)$. The distribution is zero everywhere, except for the microstate where it is infinite. Therefore the integral returns minus infinity.

No matter our definition, then, the entropy of a single microstate is minus infinity. This certainly does not sound terribly reassuring... but is it a problem?

First of all, this is in direct contradiction with the third law of thermodynamics. While the third law has many different formulations, it always leads to requiring the entropy of a perfect crystal at zero Kelvin to be zero. The idea is that a perfect crystal at zero Kelvin only has one possible configuration: it has one possible microstate. But we have seen that the entropy of a single microstate is minus infinity, not zero. Therefore the result is in contradiction.

Secondly, if we had at our disposal a system with minus infinite entropy, we could use it to circumvent the second law of thermodynamics. Suppose, in fact, we used a heat engine to transform some heat into work. By the second law, this must generate an increase in total entropy $\Delta S$. If we transfer that increase of entropy to our minus infinite entropy system, we reset the entropy of the engine and the reservoirs to the initial state. Yet this leaves the entropy of our system unchanged: minus infinity plus a finite number is still minus infinity.

Since the laws of thermodynamics are considered by some merely phenomenological, let us proceed in a different direction. Note that the entropy is additive for independent systems. Therefore the entropy of the composed system $A+B$ is $S_{A+B} = S_A + S_B$, the sum of the individual entropies. Consider now the empty system $\emptyset$. Composing any system with the empty system will have to give us back the original system $A+\emptyset = A$.\footnote{In mathematical terms, the set of all systems forms a monoid under system combination and the empty system is the identity element.} This means that the entropy of the empty system must be zero. Therefore we have a clear physical meaning for zero entropy that is, at first glance, unrelated to the third law: zero entropy is the entropy of the empty system.

Note that once we say that a system is missing, we have given a complete description of the system. There is one and only one way that a system may be missing: the empty system has only a single (discrete) microstate. Also, there is no more information that can be given about a system after one says that it is missing. An entropy lower than zero would correspond to a description that is more refined, more precise than that of an empty system. But, as we said, no such thing exists. So, from an information perspective, no system can have entropy lower than zero.\footnote{We believe this to be the true form of the third law.} A single classical microstate violates that.

So classical mechanics fails because it allows for the possibility of states that can never exist, those with minus infinite entropy. On practical grounds, we could bypass the second law with all its implications. On theoretical grounds, these ensembles do not respect the third law. On logical grounds, they would provide more information about the system than stating that the system does not exist.

We want to stress here that these entropic arguments are not independent from the isolation arguments. Thermodynamics as a physical theory is exactly about describing what happens across the boundary. Heat and work are energy exchanges by one system to others. Perfect control of a system is effectively Maxwell’s demon in disguise.

Those who are familiar with communication theory are aware that the channel capacity is given by the signal-to-noise ratio. In physics terms, the noise is the background interaction of the system with the environment. So that fixes the zero: we cannot use the system to encode more information than the background interaction allows. Saying that we can reduce the interaction with the system as much as we want effectively means setting the noise level to minus infinity, removing the background, perfectly isolating the system. In this light, all of these arguments are actually the same argument: we cannot fully isolate a system, we cannot encode infinite information in a state, we cannot have an infinite entropy differential.

\section{Quantum resolutions and open problems}

While this paper is not about providing a clear and complete understanding of quantum mechanics, we want to give a sense of how the new theory partially solves some of the problems of the old one.

To start, in quantum mechanics all pure states have zero entropy, including the vacuum, the ``empty'' system. Mixed states have positive entropy. We can no longer encode infinite information within the state of a system with continuous variables. Conceptually, the accessible information of the system is quantized by setting a lower limit, a zero, that cannot be crossed. This removes all the inconsistencies with thermodynamics, statistical mechanics and information theory. Conceptually, it signifies there is always a background interaction with the environment that cannot be fully eliminated.

If there is interaction at the system/environment boundary, then the properties of the system can only be defined up to the statistical fluctuations imposed by the interaction. In fact, in quantum mechanics the probabilistic description is baked into the very notion of state, it is not something that can be removed like in classical statistical mechanics.

If the system/environment interaction at the boundary participates in the definition of the state of a system, then states cannot be understood as intrinsic properties of the system that exist before a particular type of interaction has taken place. This is exactly what happens in quantum mechanics: an observable is well defined only when the system is put in determined conditions. A measurement may change those conditions as it is, in general, a change of the system/environment interaction. Contextuality is therefore a feature, not a problem to get rid of, and we should realize it is not even an exclusive feature of quantum mechanics.

In fact, suppose we have an open flask of gas in equilibrium. This would be described by a grand canonical ensemble, identified by its temperature $T$, volume $V$ and chemical potential $\mu$. The number of particles $N$ of the gas will not be well-defined as it fluctuates, and it is described by a probability distribution. If we want a state where $N$ is well-defined, we can simply close the flask and let the system equilibrate. We will obtain a canonical ensemble determined by the same $T$ and $V$ and by the particle number $N$. The number of particles will depend on the fluctuation at the time we closed the flask, therefore we will have a certain probability to have a particular $N$. In short, we start with a state where $\mu$ is well-defined but not $N$, we interact with the system and end up in a state where $N$ is well defined and its value determined by a probability distribution. Note the parallel with quantum mechanics: we start with a state where an observable, say the energy, is well defined and we end in a state where another observable, say position, is well defined with a value determined by a probability distribution. We change context, we change type of ensemble, we change what quantities define the ensemble. The point is that quantum contextuality is not the exception, it is classical \emph{non}-contextuality that is the exception.

For these types of reasons quantum mechanics is more precise than classical mechanics: because it better captures the definition of our system. It captures the effect of a constant minimal interaction of the system  with the environment which classical mechanics simply disregards. If anything, quantum mechanics does not go far enough.

For example, while the accessible information about the system is quantized, the possible values of the observables themselves are not. That is, position and momentum are still modeled by a real quantity. This implicitly assumes that our detectors can be made arbitrarily precise, that the resolution of our equipment can always be made finer than the resolution characterized by the quantum system. But our equipment, supposedly, is made of quantum systems. Therefore, at some point, this assumption has to fail. These types of metrological requirements will ultimately need to be incorporated into a more precise theory.

Another issue is that all pure quantum states have zero entropy and therefore the entropy cannot change during evolution. This makes it impossible to characterize what happens during out-of-equilibrium transitions, when the behavior at the boundary drastically changes. Note that quantum projections are exactly of this type. The temptation is simply to include part of the environment within the system, and then trace it out. This assumes that the interaction between the system and the part of the environment is such that it can be in principle ``switched off'', as in the case of the air with the speck of dust, and therefore does not take part in the definition of the system itself. However, it is conceivable that, while out of equilibrium, the overall system temporarily reaches states where each component cannot be described as a separate independent object. This type of interaction would necessarily influence the definition of the system, and therefore cannot be simply layered on top of the individual components. This would seem out of reach of the current quantum framework.

\section{Conclusion}

We believe these arguments help shed light onto why classical mechanics fails at a conceptual level and pave the way to a better understanding of quantum mechanics. They also give insights into what other problems still need to be addressed at a fundamental level.

We close with the following general remark. Much of the work on the foundations of physics nowadays is geared toward finding the ``true'' objects and rules of nature. Even in quantum mechanics, a lot of efforts are aimed at finding the true underlying constituents and mechanisms that would explain the quantum world. In this sense, theoretical physics is built bottom-up. We believe this general approach to be problematic.

Experimental physics is built top-down: we start with big objects, and decompose them into smaller and smaller parts. The objects of our research are defined by what parts can be carved out, accessed and manipulated experimentally, independently from one another. This imposes a number of requirements, constraints and assumptions that allow us to define the physical system. These definitional requirements are, therefore, not practical considerations that only experimentalists should be bothered by. They must be at the foundation of our theoretical treatment since, without their stipulation, there is no physical system to study. In this sense, they are more fundamental than the laws the system obeys, because we cannot define the law before having a well-defined system.

We believe it is the inclusion of these definitional requirements and assumptions, which we have started to sketch here and in our more general project Assumptions of Physics, that will ultimately form a more sound foundation for physical theories.

\bibliography{bibliography}

\end{document}

### 2021-08-13 Overviews/Differentiability/Differentiability.tex
Differentiability: added more ideas

@@
 
+\subsection{Deep approaches}
+
+These approaches try to formulate differentiability in terms of physical significance. The status of differentiable coordinates is that they are needed to validly express the physics problem.
+
+\subsubsection{Smoothness as coordinates with continuous Radon-Nikodym derivate}
+
 \begin{defn}
@@
 
+The idea is that continuity of the derivative is needed to express the dependence of the density from the coordinates in a meaningful way (from previous work, only continuous function can expressed experimentally meaningful quantities).
+
+The issue is that, in general, a single measure is not going to be enough: regions where the measure is zero would allow for non-differentiable coordinates (the derivative would be zero, which is always continuous).
+
+More physical would be to consider the space of all measures, which would correspond to all possible ways a distributive quantity can be distributed (i.e.
+the states of a composite system).
+This, however, does move the problem of differentiability to the space of measures.
+In general, the derivative will be measurable and not continuous.
+Therefore one would need a way to constrain the space to only the ones that are continuous.
+
+General issue: more thought needs to be put into how densities are actually measured, and what would be the correct topology on the space.
+
+\subsection{Shallow approaches}
+
+These approaches do not try to assign physical significance to differentiability. The status of differentiable coordinates is that they are mathematically convenient.
+
+\subsubsection{Smoothness as convenient to express approximations of linear functionals}
+
+\begin{defn}
+	Let $X$ be a manifold of dimension $n$.
+A \textbf{k-surface} is a subset $U \subseteq X$ that is a $k$-dimensional manifold (with boundary) with the subspace topology.
+We note $S^k(X)$, or simply $S^k$, the set of all $k$-surfaces and $S(X) = \bigcup_{i=0}^{n} S^k$ the set of all possible $k$-surfaces of all dimensions.
+\end{defn}
+
+\begin{remark}
+	Note that $S^k = \emptyset$ if $k < 0$ or $k > n$.
+\end{remark}
+
+\begin{defn}
+	The \textbf{boundary operator} $\partial : S(X) \to S(X)$ is a map from a $k$-surface to its boundary (as a manifold).
+\end{defn}
+
+\begin{prop}
+	For every $\sigma^k \in S^k(X)$ with $k > 0$, we have $\partial \sigma^k \in S^{k-1}$.
+For every $\sigma^0 \in S^0$, $\partial \sigma^0 = \emptyset$.
+For every $\sigma \in S$, we have $\partial \partial \sigma = \emptyset$.
+\end{prop}
+
+\begin{defn}
+	A \textbf{k-functional} is a function $f_k : S^k \to \mathbb{R}$ with the following properties:
+	\begin{description}
+		\item[Linear] $f_k(\sigma_1 \cup \sigma_2) = f_k(\sigma_1) + f_k(\sigma_2)$ for every $\sigma_1, \sigma_2 \in S^k$ such that $\sigma_1 \cap \sigma_2 = \emptyset$
+		\item[No contribution from boundary] $f_k(\sigma) = f_k(\sigma \setminus \partial \sigma)$ for every $\sigma \in S^k$
+		\item[Commutes with the limit] $\lim\limits_{i \to \infty} f_k(\sigma_i^k) = f_k(\lim\limits_{i \to \infty}\sigma_i^k)$
+	\end{description}
+\end{defn}
+
+\begin{remark}
+	Linearity means that the contribution from each surface is independent from the rest (reducibility).
+If boundaries gave contribution, the functional would not really be a function of the $k$-surface, since the boundary is of lower dimension.
+
+	Note the operator $length : S^1 \to \mathbb{R}$ that returns the length of a line is not a $1$-functional as it does not commute with the limit (e.g.
+diagonal is the limit of a perpendicular zig-zag, but the length does not converge).
+\end{remark}
+
+\begin{desid}
+	Let $X$ be a differentiable manifold.
+For every $k$-functional $f_k$ there exists a suitable $k$-form $\omega_k$ whose integral $\int_U \omega_k$ in a region $U$ approximates $f_k(U)$ at a desired level of precision.
+\end{desid}
+
+\begin{remark}
+	The simplest strategy may be to fix an interval for the value, and use standard criteria of convergence. However, this may be far too restrictive with respect to what we need physically.
+
+	Given that we measure the quantity $f_k(\sigma^k)$ in a finite region $\sigma_k$ with finite precision, there are two things to fix: the confidence interval $\epsilon \in \mathbb{R}$ for the quantity itself and the smallest regions we can measure.
+As we impose a cutoff in the size of the regions, we do not care whether the approximation holds well at smaller scale (e.g.
+anything ``higher frequency'' can be ignored, and maybe it doesn't even matter whether higher frequencies diverge).
+
+	We want to define those without reference to a metric/measure on $X$.
+\end{remark}
+
+\begin{defn}
+	A \textbf{coarsening} is a map $c : S^k \to S^k$ with the following properties:
+	\begin{description}
+		\item[monotonic] $c(\sigma) \supseteq \sigma$ for all $\sigma \in S^k$
+		\item[idempotent] $c(c(\sigma)) = c(\sigma$) for all $\sigma \in S^k$
+		\item[bounded] for all $\sigma \in S^k$, there exists a $\hat{\sigma} \in S^k$ such that $\hat{\sigma} \subseteq \sigma$ and $c(\hat{\sigma}) \neq \hat{\sigma}$
+	\end{description}
+	A \textbf{coarsened region} is a $k$-surface $\sigma_k$ such that $c(\sigma_k) = \sigma_k$. 
+\end{defn}
+
+\begin{remark}
+	The idea is that the coarsening substitutes the measuring region to one bigger.
+Monotony guarantees that the region is bigger.
+Boundedness guarantees that there is a cutoff in size at a certain point.
+Idempotency guarantees that there is a desired resolution.
+
+	A simple example of coarsening: consider the set of all surfaces with integer areas in some coordinates; given any surface, grow it to match one of those.
+\end{remark}
+
+\begin{desid}
+	Given a $k$-functional $f_k : S^k \to \mathbb{R}$, a coarsening $c : S^k \to S^k$ and an $\epsilon \in \mathbb{R}$, we can always find a $k$-form $\omega_k$ such that $| f_k(\sigma) - \int_\sigma \omega_k | < \epsilon$ for all coarsened regions $\sigma$.
+\end{desid}
+
 %\section{Examples}

### 2021-08-13 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
ClassicalFail: adding todos for citations

@@
 The common view is that classical mechanics is a perfectly consistent theory that just does not happen to be true.
-It fails simply because experiments do not agree with it.
-This view is so established, so ingrained, that it is seldom discussed explicitly, and we believe it to be part of the reason why quantum mechanics still feels, after almost a century, so mysterious: how are we going to conceptually appreciate a theory if we do not even know what conceptual problems it is trying to address?
+It fails simply because experiments do not agree with it.\footnote{CITE: papers that either make the claim, or that discuss how everybody makes that claim} This view is so established, so ingrained, that it is seldom discussed explicitly, and we believe it to be part of the reason why quantum mechanics still feels, after almost a century, so mysterious: how are we going to conceptually appreciate a theory if we do not even know what conceptual problems it is trying to address?
 
@@
 Conceptually, we can think of a property as being intrinsic to the system only if the interaction with the environment leaves it, at least to a good approximation, undisturbed.
-If it is not left undisturbed, then it is no longer a property of the system itself: it is a shared property between the system and the environment because the nature of both comes into play in its definition.
+If it is not left undisturbed, then it is no longer a property of the system itself: it is a shared property between the system and the environment because the nature of both comes into play in its definition.\footnote{CITE: maybe someone in complex systems, emergent properties, ...
+make similar claims}
 
@@
 
-Apart from practical concerns, any wall we will use as shielding will necessarily have finite temperature and will emit thermal radiation; moreover, gravitational interactions cannot be shielded in the first place.
-Even theoretically, complete isolation is not possible.
+Apart from practical concerns, any wall we will use as shielding will necessarily have finite temperature and will emit thermal radiation; moreover, gravitational interactions cannot be shielded in the first place.\footnote{CITE: something that goes more into the details as to why gravitational interactions cannot be shielded} Even theoretically, complete isolation is not possible.
 We cannot, even conceptually, instantly remove all the background interactions.
@@
 
-Since this is a point of contention, and we do not believe further discussion along these lines will change anyone's perspective, let us switch gears.
+Since this is a point of contention,\footnote{CITE: papers that show different view, papers that claim there are different views.} and we do not believe further discussion along these lines will change anyone's perspective, let us switch gears.
 We are going to argue that a perfect description of the type provided by classical mechanics is inconsistent with thermodynamics and statistical mechanics.
@@
 We start by asking: what is the entropy of a single microstate?
-This is complicated by the fact that we have (at least) two definitions of entropies, and people argue about which is the correct one.
-Therefore let us use both definitions and see what happens.
+This is complicated by the fact that we have (at least) two definitions of entropies, and people argue about which is the correct one.\footnote{CITE: papers with different stances on entropy; papers that claim different stances} Therefore let us use both definitions and see what happens.
 
@@
 
-Since the laws of thermodynamics are considered by some merely phenomenological, let us proceed in a different direction.
+Since the laws of thermodynamics are considered by some merely phenomenological,\footnote{CITE: papers that claim that thermodynamic laws are only valid on average} let us proceed in a different direction.
 Note that the entropy is additive for independent systems.
@@
 
-Those who are familiar with communication theory are aware that the channel capacity is given by the signal-to-noise ratio.
-In physics terms, the noise is the background interaction of the system with the environment.
+Those who are familiar with communication theory are aware that the channel capacity is given by the signal-to-noise ratio.\footnote{CITE: papers that define channel capacity (Shannon?)} In physics terms, the noise is the background interaction of the system with the environment.
 So that fixes the zero: we cannot use the system to encode more information than the background interaction allows.
@@
 Much of the work on the foundations of physics nowadays is geared toward finding the ``true'' objects and rules of nature.
-Even in quantum mechanics, a lot of efforts are aimed at finding the true underlying constituents and mechanisms that would explain the quantum world.
-In this sense, theoretical physics is built bottom-up.
+\footnote{CITE: papers that propose emergent models of space time, ...} Even in quantum mechanics, a lot of efforts are aimed at finding the true underlying constituents and mechanisms that would explain the quantum world.\footnote{CITE: papers that propose underlying mechanism (e.g.
+Bohmian approach, GRW)} In this sense, theoretical physics is built bottom-up.
 We believe this general approach to be problematic.

### 2021-08-24 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
ClassicalFail: adding a couple of references

@@
 We close with the following general remark.
-Much of the work on the foundations of physics nowadays is geared toward finding the ``true'' objects and rules of nature.
-\footnote{CITE: papers that propose emergent models of space time, ...} Even in quantum mechanics, a lot of efforts are aimed at finding the true underlying constituents and mechanisms that would explain the quantum world.\footnote{CITE: papers that propose underlying mechanism (e.g.
+Much of the work on the foundations of physics nowadays is geared toward finding the ``true'' objects and rules of nature.\cite{huggett2013emergent,oriti2009approaches,lee2020model} \footnote{CITE: papers that propose emergent models of space time, ...} Even in quantum mechanics, a lot of efforts are aimed at finding the true underlying constituents and mechanisms that would explain the quantum world.\cite{larsson2014loopholes, ghirardi1986unified, carroll2019something, nikolic2007bohmian}\footnote{CITE: papers that propose underlying mechanism (e.g.
 Bohmian approach, GRW)} In this sense, theoretical physics is built bottom-up.

### 2021-08-24 Overviews/Differentiability/Differentiability.tex
Differentiability: clarifying continuity of functional

@@
 
+\subsection{General definitions}
+
+\begin{defn}
+	Let $X$ be a manifold of dimension $n$.
+A \textbf{k-surface} is a subset $U \subseteq X$ that is a $k$-dimensional manifold (with boundary) with the subspace topology (i.e.
+regular/embedded submanifold).
+We note $S^k(X)$, or simply $S^k$, the set of all $k$-surfaces and $S(X) = \bigcup_{i=0}^{n} S^k$ the set of all possible $k$-surfaces of all dimensions.
+\end{defn}
+
+\begin{remark}
+	Note that $S^k = \emptyset$ if $k < 0$ or $k > n$.
+\end{remark}
+
+\begin{defn}
+	The \textbf{boundary operator} $\partial : S(X) \to S(X)$ is a map from a $k$-surface to its boundary (as a manifold).
+\end{defn}
+
+\begin{prop}
+	For every $\sigma^k \in S^k(X)$ with $k > 0$, we have $\partial \sigma^k \in S^{k-1}$.
+For every $\sigma^0 \in S^0$, $\partial \sigma^0 = \emptyset$.
+For every $\sigma \in S$, we have $\partial \partial \sigma = \emptyset$.
+\end{prop}
+
+\begin{defn}
+	A \textbf{k-functional} is a function $f_k : S^k \to \mathbb{R}$ with the following properties:
+	\begin{description}
+		\item[Linear] $f_k(\sigma_1 \cup \sigma_2) = f_k(\sigma_1) + f_k(\sigma_2)$ for every $\sigma_1, \sigma_2 \in S^k$ such that $\sigma_1 \cap \sigma_2 = \emptyset$
+		\item[No contribution from boundary] $f_k(\sigma) = f_k(\sigma \setminus \partial \sigma)$ for every $\sigma \in S^k$
+		\item[Commutes with the limit] $\lim\limits_{i \to \infty} f_k(\sigma_i^k) = f_k(\lim\limits_{i \to \infty}\sigma_i^k)$ (limit of a sequence of submain)
+	\end{description}
+\end{defn}
+
+\begin{remark}
+	Linearity means that the contribution from each surface is independent from the rest (reducibility).
+If boundaries gave contribution, the functional would not really be a function of the $k$-surface, since the boundary is of lower dimension.
+
+	Note the operator $length : S^1 \to \mathbb{R}$ that returns the length of a line is not a $1$-functional as it does not commute with the limit (e.g.
+diagonal is the limit of a perpendicular zig-zag, but the length does not converge).
+\end{remark}
+
+\subsubsection{Continuity for functionals}
+
+The requirements of commutation with the limit is similar to a definition of continuity in terms of limits.
+Can we turn that into a topological notion of continuity?
+Suppose we had a topology on $S^k$, could we simply require the functional to be continuous?
+How that would constrain $f_k$?
+
+This forms a nice parallel between a map $f : X \to Y$ and the functional $f_k : S^k \to \mathbb{R}$.
+For both, the open sets are associated to the measurement precision.
+For $f_k$, the precision of the quantity is an open interval (finite precision) while the precision of the argument is a set of possible $k$ surfaces (i.e.
+we have an uncertainty on the shape/area).
+
+For example, let $U, V \in S^k$ be two $k$-surfaces without borders such that $U \subset V$.
+The set $R(U,V) = \{ W \in S^k \, | \, U \subseteq W \subseteq V\}$ is called the \textbf{surface range} between $U$ and $V$.
+That should be an open set in the topology of $S^k$.
+
 \subsection{Deep approaches}
@@
 
+\subsubsection{title}
+
 \subsection{Shallow approaches}
@@
 
-\subsubsection{Smoothness as convenient to express approximations of linear functionals}
+\subsubsection{Convergence at scale}
+
+We need to understand what is the proper criteria for convergence that correctly reflect the physical approximations we are looking for.
+The general idea is that we want an approximation that works from a given level of scale upwards, leaving errors at lower scale.
+This is not how typically things work with standard types of convergence.
+
+Suppose we have a function $f(x)$.
+Consider now $\hat{f}(x) = f(x) + \epsilon$, where $\epsilon$ can be understood as a small constant error over the whole space.
+The difference in integral $\int_U \hat{f}(x) - \int_U f(x)dx = \epsilon \Delta x$ increases with the size of the region.
+In the case the values $x$ and $f(x)$ represent directly the value measured (e.g.
+$x$ is temperature and $f(x)$ is the length of a metal bar), $\hat{f}$ represents a good approximation.
+If what we measure directly, however, is the integral (e.g.
+$U$ is a region and $\int_U f(x)dx$ is the mass in that region), then $\hat{f}$ would give a poor approximation at large scale.
+For this second case, we need a different method of convergence.
+
+A first attempt could be the following.
 
 \begin{defn}
-	Let $X$ be a manifold of dimension $n$.
-A \textbf{k-surface} is a subset $U \subseteq X$ that is a $k$-dimensional manifold (with boundary) with the subspace topology.
-We note $S^k(X)$, or simply $S^k$, the set of all $k$-surfaces and $S(X) = \bigcup_{i=0}^{n} S^k$ the set of all possible $k$-surfaces of all dimensions.
+	Let $X$ be a differentiable manifold.
+Let $d : X \times X \to [0, +\infty)$ be a metric that induces the topology of $X$.
+Let $B^k(X, \delta)$ be the set of all $k$-dimensional balls with radius greater than $\delta$.
+Let $\omega$ be a $k$-form.
+A sequence of $k$-forms $\{\omega_i\}$ \textbf{converges at scale} to $\omega$ if given $\epsilon, \delta \in \mathbb{R}^+$ we can find $N$ such that $\| \int_U \omega_n - \int_U \omega \| < \epsilon$ for all $n \geq N$ and $U \in B^k(X, \delta)$.
 \end{defn}
 
-\begin{remark}
-	Note that $S^k = \emptyset$ if $k < 0$ or $k > n$.
-\end{remark}
-
 \begin{defn}
-	The \textbf{boundary operator} $\partial : S(X) \to S(X)$ is a map from a $k$-surface to its boundary (as a manifold).
+	Let $X$ be a differentiable manifold.
+Let $d : X \times X \to [0, +\infty)$ be a metric that induces the topology of $X$.
+Let $B^k(X, \delta)$ be the set of all $k$-dimensional balls with radius greater than $\delta$.
+Let $f_k$ be a $k$-functional.
+A sequence of $k$-forms $\{\omega_i\}$ \textbf{converges at scale} to $f_k$ if given $\epsilon, \delta \in \mathbb{R}^+$ we can find $N$ such that $\| \int_U \omega_n - f_k(U) \| < \epsilon$ for all $n \geq N$ and $U \in B^k(X, \delta)$.
 \end{defn}
 
-\begin{prop}
-	For every $\sigma^k \in S^k(X)$ with $k > 0$, we have $\partial \sigma^k \in S^{k-1}$.
-For every $\sigma^0 \in S^0$, $\partial \sigma^0 = \emptyset$.
-For every $\sigma \in S$, we have $\partial \partial \sigma = \emptyset$.
-\end{prop}
-
-\begin{defn}
-	A \textbf{k-functional} is a function $f_k : S^k \to \mathbb{R}$ with the following properties:
-	\begin{description}
-		\item[Linear] $f_k(\sigma_1 \cup \sigma_2) = f_k(\sigma_1) + f_k(\sigma_2)$ for every $\sigma_1, \sigma_2 \in S^k$ such that $\sigma_1 \cap \sigma_2 = \emptyset$
-		\item[No contribution from boundary] $f_k(\sigma) = f_k(\sigma \setminus \partial \sigma)$ for every $\sigma \in S^k$
-		\item[Commutes with the limit] $\lim\limits_{i \to \infty} f_k(\sigma_i^k) = f_k(\lim\limits_{i \to \infty}\sigma_i^k)$
-	\end{description}
-\end{defn}
-
-\begin{remark}
-	Linearity means that the contribution from each surface is independent from the rest (reducibility).
-If boundaries gave contribution, the functional would not really be a function of the $k$-surface, since the boundary is of lower dimension.
-
-	Note the operator $length : S^1 \to \mathbb{R}$ that returns the length of a line is not a $1$-functional as it does not commute with the limit (e.g.
-diagonal is the limit of a perpendicular zig-zag, but the length does not converge).
-\end{remark}
+Then we would like to prove the following:
 
 \begin{desid}
-	Let $X$ be a differentiable manifold.
-For every $k$-functional $f_k$ there exists a suitable $k$-form $\omega_k$ whose integral $\int_U \omega_k$ in a region $U$ approximates $f_k(U)$ at a desired level of precision.
+	Let $X$ be a differentiable manifold. For every $k$-functional $f_k$ there exists a sequence of $k$-forms $\{\omega_i$\} that converges at scale to $f_k$.
 \end{desid}
 
-\begin{remark}
-	The simplest strategy may be to fix an interval for the value, and use standard criteria of convergence. However, this may be far too restrictive with respect to what we need physically.
+\section{Graveyard}
 
-	Given that we measure the quantity $f_k(\sigma^k)$ in a finite region $\sigma_k$ with finite precision, there are two things to fix: the confidence interval $\epsilon \in \mathbb{R}$ for the quantity itself and the smallest regions we can measure.
-As we impose a cutoff in the size of the regions, we do not care whether the approximation holds well at smaller scale (e.g.
-anything ``higher frequency'' can be ignored, and maybe it doesn't even matter whether higher frequencies diverge).
+\subsection{Coarsening}
 
-	We want to define those without reference to a metric/measure on $X$.
-\end{remark}
+This was an attempt to define how to coarsen integrand regions in a way that doesn't depend on a metric.
 
@@
 
-%\section{Insightful failures}
-
 \bibliography{bibliography}

### 2021-10-10 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: adding comments after review

@@
 All the propositions of quantum logic are consequently readily found in the Borel algebra of the Hilbert space.
-The Borel algebra, in turn, is a Boolean algebra that follows classical logic, and contains all the statements of quantum logic.
-This construction also applies identically for the classical cases, where we will also be able to talk about the set of all closed subspaces.
+The Borel algebra, in turn, is a Boolean algebra that follows classical logic, and contains all the statements of quantum logic.\footnote{Obviously, the inclusion map from the lattice of closed subspaces is not an order isomorphism as the join and complement are not preserved.
+However, these are still expressible as subspace closure and subspace complement.} This construction also applies identically for the classical cases, where we will also be able to talk about the set of all closed subspaces.
 
@@
 Renormalization would lead to the set $\{ \psi \in \mathcal{H} \, | \, \langle \psi , A \psi \rangle / \langle \psi , \psi \rangle \in (x_0, x_1) \}$, which is still a Borel set since renormalization is a continuous function over its domain $\mathcal{H} \setminus \{ 0 \}$.} Consider, for example, ``the expectation of $z$-spin is zero''.
-This, by Proposition 2, would correspond to a Borel set $V \subseteq \mathcal{H}$.
-This set is not the full set since the state with $z$-spin up is not in the set.
-However, $x$-spin up and $x$-spin down will be in the set, and their span is the full space.
+This would correspond to the set $V \subseteq \mathcal{H}$ that includes all states, and only the states, for which the expection of $S_z$ is zero.
+By Proposition 2, $V$ is a Borel set.
+In fact, the set $U = {0}$ is a Borel set of $\mathbb{R}$ and therefore $V = F_{S_z}^{-1}(U)$ is a Borel set of $\mathcal{H}$.
+However, $V$ is not a closed subspace.
+The set $V$ is not the full set $\mathcal{H}$ since, for example, the state with $z$-spin up is not in the set.
+However, $x$-spin up and $x$-spin down will be in the set $V$, and their span is the full space $\mathcal{H}$.
 Therefore $V$ is not a closed subspace.
-Our proposition, despite being physically interesting, does not correspond to a closed subspace.
+Our proposition, despite being physically interesting, does not correspond to a closed subspace but does correspond to a Borel set.
 
@@
 In fact, no cross section measurement, our main tool in exploring the standard model, is single-shot.
-Being that quantum mechanics is a statistical theory, most physically interesting statements are statistical in nature, and that is why we need to include the Borel sets in our lattice of statements.
+Being that quantum mechanics is a statistical theory, most physically interesting statements are statistical in nature, and that is why we need to include the Borel sets in our lattice of statements.\footnote{Since probabilities are assigned to Borel sets and only Borel sets, the Borel algebra will contain all physically interesting statements.
+Further expansions (e.g.
+to the power set) would only include unphysical statements.}
 

### 2021-10-10 Overviews/Experimental verifiability/ExperimentalVerifiabilityOverview.tex
Overviews: experimental verifiability review

@@
-\documentclass[10pt,twocolumn, nofootinbib]{revtex4-1}
+\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
 %\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
@@
 
-\input{include/basicstyle}
-\input{include/logic}
-\input{include/theorems}
-
+\usepackage{assumptionsofphysics}
 \usepackage{graphicx}
@@
 
-\title{Assumptions of Physics overview: \\
-	experimental verifiability and topology }
+\title{Experimental verifiability and topology }
 \author{Gabriele Carcassi, Christine A. Aidala}
@@
 These insights can be taken as a first step in a general mathematical theory for experimental science.
-This work is an overview of some of the results of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
 \end{abstract}
@@
 
-The overall structure (see \cite{AoPBook} for more details) can be summed up in the following diagram that can be used as a guide throughout this note.
+The overall structure (see \cite{aop-book} for more details) can be summed up in the following diagram that can be used as a guide throughout this note.
 

### 2021-10-10 Overviews/Quantum mechanics/QuantumMechanicsOverview.tex
Overviews: updating revtex4 reference
Overviews: final pass to quantum mechanics

@@
 \begin{abstract}
-We briefly show how quantum mechanics can be rederived and better understood as a consequence of irreducibility.
+We briefly summarize our strategy to understand and rederive quantum mechanics as a consequence of irreducibility.
 The Schroedinger equation (i.e.
 unitary evolution) is recovered as deterministic and reversible motion while the measurement process is recovered as a special case of a process with equilibria (i.e.
-idempotent process, projection).
-This work is an overview of some of the results of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
+idempotent linear process, projection).
 \end{abstract}
@@
 
-The overall argument (see \cite{AoPPhy1,AoPBook} for more details) can be summed up in the following diagram that can be used as a guide throughout this note.
+The ideas and result summarized here are part of our larger project Assumptions of Physics (see \cite{AoPPhy1,AoPBook} for more details).
+They represent the current best understanding and will be revised as the project matures.
+
+This note can be summed up in the following diagram that can be used as a guide throughout this note.
 
@@
 The state of a fragment will be associated with a random variable uniformly distributed over the possible fragment states.
-As discussed in the context of classical mechanics(CITE), distributions over states must be invariant and symplectic manifolds are the only manifolds over which invariant distributions can be defined.
+As we show in the context of classical mechanics, distributions over states must be invariant and symplectic manifolds are the only manifolds over which invariant distributions can be defined.
 As we cannot say anything about the state of the fragments, the dimensionality of this manifold must be irrelevant as long as it is even dimensional.
 For simplicity, we can choose a two-dimensional one.
-Therefore we are interested in the space of bi-dimensional uniform distributions formed by a pair of two random variables $A$ and $B$.
+Therefore we are interested in the space of bi-dimensional uniform distributions formed by a pair of two random variables $A$ and $B$.\footnote{We are looking into alternative strategies for this part, as we are not fully satisfied with it.}
 

### 2021-10-10 Overviews/Classical mechanics/ClassicalMechanicsOverview.tex
Overviews: updating revtex4 reference
Overviews: classical mechanics review

@@
 
-\title{Assumptions of Physics overview: \\
-	Classical mechanics and infinitesimal reducibility }
+\title{Classical mechanics and infinitesimal reducibility }
 \author{Gabriele Carcassi, Christine A. Aidala}
@@
 We briefly show how classical mechanics can be rederived and better understood as a consequence of three assumptions: infinitesimal reducibility, deterministic and reversible evolution, and kinematic equivalence.
-This work is an overview of some of the results of Assumptions of Physics, a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
 \end{abstract}
@@
 
-The overall argument (see \cite{AoPPhy1,AoPBook} for more details) can be summed up in the following diagram that can be used as a guide throughout this note.
+The overall argument (see \cite{AoPPhy1,aop-book} for more details) can be summed up in the following diagram that can be used as a guide throughout this note.
 

### 2021-10-10 Articles/2021-NatureLetter/quantum_state_final.tex
PBRLetter: submitted version

%\documentclass[11pt, executivepaper]{article}
%\usepackage[utf8]{inputenc}
%\usepackage[T1]{fontenc}
%\usepackage{natbib}
%\usepackage{amsmath}
%\usepackage{mathtools}
%\usepackage{xcolor}
%\usepackage{amsfonts}
%\usepackage{graphicx}
%\usepackage{enumitem}

\documentclass[aps,twocolumn,showpacs,preprintnumbers]{revtex4}
\usepackage{graphicx}  % Include figure files
\usepackage{subfigure}
\usepackage{multirow}
\linespread{1.1}
\usepackage{fancyhdr}
\usepackage{longtable}
\usepackage{parskip}
\usepackage[T1]{fontenc}
\usepackage{dcolumn}
\usepackage{bm}        
\usepackage{amsfonts}  
\usepackage{amsmath}   
\usepackage{amssymb}
\usepackage{hyperref}
\hypersetup{colorlinks= true, allcolors=blue}
\setcitestyle{aysep={}}
\setlength{\parindent}{10pt}

\begin{document}

	\title{\textbf{On the Reality of the Quantum State Once Again (TODO)}}

	\author{Andrea Oldofredi, Gabriele Carcassi}
	\affiliation {\it Universit\'e de Lausanne, Switzerland, University of Michigan, Ann Arbor}

	\date{September 4, 2021}

	\maketitle

	In 2012 Pusey, Barrett and Rudolph published a formal result in \emph{Nature Physics}, widely known as the PBR theorem, showing that ``if the quantum state merely represents information about the real physical state of a system, then experimental predictions are obtained that contradict those of quantum theory'' (\cite{PBR:2012}, p.\ 475). Alternatively stated, PBR argued that in every model reproducing the statistics and predictions of Quantum Mechanics (QM) the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi-$ontic. Consequently, quantum theories cannot be $\psi-$epistemic.

	Such a theorem has created controversies on its actual meaning: on the one hand, PBR claim that it rules out epistemic interpretations of quantum theory according to which quantum systems do have physical states, but $\psi$ merely represents information---as for instance in Rovelli's Relational Quantum Mechanics \cite{Rovelli:1996}. On the other hand, many scholars have recently shown that epistemic as well as statistical approaches to quantum theory are not refuted by the PBR argument (cf.\ \cite{Ben:2017}, \cite{Rizzi:2018}, \cite{Oldofredi:2021}, \cite{DeBrota:2019}).

	In this letter we are not going to rebut the theorem itself. We instead aim to carefully re-examine its premises, highlighting an implicit assumption which we believe forces us to re-evaluate the significance of the conclusion. (REVISED)

	The structure of their argument rests on a few explicit premises: (i) isolated quantum systems have real, objective, observer-independent states, (ii) systems prepared independently have independent physical states, and (iii) the latter are correctly described by QM. By stipulating that $\psi$ represents only agent's uncertainty about the actual state of the prepared systems, PBR obtain a contradiction with their initial assumptions concluding that quantum theories cannot be $\psi-$epistemic.

	Looking at their proof, however, there is a fourth implicit hypothesis: (iv) during a measurement the real state of the system under consideration remains unaltered by the observation. This means that such physical state does not change during the measurement process. Indeed, the authors' argument requires that initially prepared states should be found after the observation. This fact is inherited from Harrigan and Spekkens' classification between $\psi-$ontic and $\psi-$epistemic models (cf.\ \cite{Harrigan:2010}), on top of which the PBR theorem is derived.

	According to this categorization, which is framed within operational QM, ``the primitives of description are the properties of microscopic systems. A preparation procedure is assumed to prepare a system with certain properties and a measurement procedure is assumed to reveal something about those properties. A complete specification of the properties of a system is referred to as the ontic state of that system'' (\cite{Harrigan:2010}, p.\ 128).

	Now, one may rephrase the theorem by saying that any model in which a measurement represents a mere update of information with no effect on the underlying physical state, must make predictions which contradict those of quantum theory. In fact, since in general the quantum state \emph{does} change during a measurement, in a model in which the quantum state represents the ontic state of the system, $\psi$ must change as well. Similarly a model in which the state represents information, but for which the physical state changes during a measurement is not ruled out either, because the initial information becomes obsolete and the measurement process will update it. Thus, we may restate the conclusion of the PBR theorem more precisely as follows: ``\emph{any} quantum model in which a measurement leaves the physical state unchanged must make predictions which contradict those of QM''.

	Furthermore, it is useful to underline that Harrigan and Spekkens' statement quoted above---and consequently assumption (i) (TODO: is this i or iv ? )---is generally not correct in the context of standard QM for two reasons: (i) quantum measurements usually do not reveal pre-existing values of properties, and (ii) the ontic state of the system in non-measurement situations is strongly undetermined (i.e.\ the properties of a non-observed system have non-definite values). Hence, it seems that the PBR theorem rests on an ill-defined basis, given that in every empirically adequate quantum model, no matter if ontic or epistemic, measurements do alter the states of the quantum systems (epistemic interpretations in which there are no physical objects such as QBism avoid this issue by construction).

	Moreover, we can reinterpret this result as a non-contextuality claim. We may rephrase this theorem by saying that models in which quantum measurements do not produce any change to the observed quantum systems contradict the predictions and statistics of QM. It follows that it can be considered another no-go theorem refuting the non-contextuality of quantum observables, along the lines of the Kochen-Specker and Gleason's theorems (cf.\ \cite{Kochen:1967}, \cite{Gleason:1957}).

	In conclusion, the PBR theorem does not rule out all models in which $\psi$ represents information, only those in which the physical state is assumed not to change during a measurement. This should be however hardly surprising given other well-established results in the field of quantum foundations.

	%It is not clear to us whether the intent of the authors was to rule out specifically those models for which the state does not change during a measurement or not. It is not clear to us whether the distinction is clear to most physicists. We should point out that there are already measurable effects that show that the measurement does in fact produce physical results. Most of all, the Quantum Zeno effect (also known as "watched pot never boils") in which repeated frequent observation keeps collapsing the state of the system to the original state, inhibiting the natural evolution. This effect has been measured in different settings in accordance to theoretical predictions. Though we find the technique interesting, we don't see that the theorem adds much, in the sense that if the Quantum Zeno effect is not proof enough to make one conceed that measurements are processes that affect the physical state of the system, we do not see what will.

	%So, what is the point of all this discussion? The PBR theorem does not rule out all models in which the state is information, only the ones where the physical state is assumed to not change during a measurement. This should be hardly surprising giving other predicted and measured effects, like the Quantum Zeno effect. While information theory cares little about what information needs to be kept in a specific case and what causes the information to change. Characterizing, implicitly or explicitly, the correct set of variables to be used for a specific class of systems and processes is the actual physics problem.

	\bibliographystyle{ieeetr}
	\bibliography{PhDthesis}
\end{document}

### 2021-10-18 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QLogic: Andrea's modifications after first review

@@
 
-As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus, a belief that is widespread still to this day (cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004} and \cite{deRonde:2016}).
-From the thirties, indeed, various modifications of Classical Logic (CL) were advanced in order to faithfully represent the physical content of quantum theory.
+As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus.\footnote{For an introduction and a clear discussion of quantum logic the reader may refer to cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{deRonde:2016}, \cite{Wilce:2021}.} This belief stood the test of time, and it is still widespread to the present day.
+For example, \cite{Giuntini:2018} claim that ``[t]he \emph{logic} of classical physical objects is naturally based on a two-valued semantics.
+Such a dichotomic situation breaks down in quantum theory'' (p.\ 4), implying that the latter entails a revision of classical propositional calculus.
+Such a view is shared by \cite{Wilce:2021}, who says: ``[m]athematically, quantum mechanics can be regarded as a non-classical probability calculus resting upon a non-classical propositional logic''.
+\noindent Similarly, presenting the standard formulation of quantum theory and its peculiar features, Stefanovich states that
+\begin{quote}
+the true logical relationships between results of measurements are different from the classical laws of Aristotle and Boole.
+The usual classical logic needs to be replaced by the so-called quantum logic (\cite{Stefanovich:2019}, p.\ 2)
+\end{quote}
+\noindent Even more explicitly, introducing the quantum logical formalism Holik writes that
+\begin{quote}
+[t]he connectives ``$\vee$'', ``$\wedge$'' and ``$\neg$'' are the quantum logical analogues of the Boolean connectives.
+It is important to remark that, with these connectives, the lattice of closed subspaces does not satisfies [\emph{sic}] the distributivity law [...] and then, it is non-Boolean.
+This implies that the new connectives cannot be interpreted in a classical way (\cite{Holik:2021}, p.\ 5).
+\end{quote}
+
+The first modifications of Classical Logic (CL) were advanced in the thirties in order to faithfully represent the physical content of quantum theory.
 For instance, in 1931 the Polish philosopher Zygmunt Zawirski proposed to apply \L ukasiewicz's three-valued logic to QM, starting from considerations about the wave-particle duality and Heisenberg's uncertainty principle.
@@
 
-A few years after the publication of this paper, the research in quantum logic saw a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct logical structures to understand and describe the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}).
+A few years after the publication of this paper, the research in quantum logic saw a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct logical structures to understand and describe the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{Pitowsky:2006}, \cite{Giuntini:2018}, \cite{Svozil:2020}, \cite{Holik:2021}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}).
 In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography.
@@
 In this paper we will be concerned with a similar issue, since we ask whether standard quantum mechanics itself necessarily involves a new logical structure radically different with respect to classical propositional calculus.
-In what follows we will answer this question in the negative.
-More precisely, this essay has a two-fold aim: in the first place, we analyze the main motivations and justifications for the introduction of quantum logic, claiming that such arguments do not really entail the failure of CL in the quantum context because they contain serious flaws.
+In what follows we will answer this question in the negative.\footnote{It is worth noting that in what follows we are \emph{not} going to discuss whether it is useful or possible to describe quantum phenomena with a non-classical type of logic.} More precisely, this essay has a two-fold aim: in the first place, we analyze the main motivations and justifications for the introduction of quantum logic, claiming that such arguments do not really entail the failure of CL in the quantum context because they contain serious flaws.
 In the second place, we show that the mathematical structure on which QM is built already provides us with a distributive lattice of quantum propositions, contrary to the received view in QL.
@@
 
-The idea is that the logical operations of conjunction, disjunction and negation are replaced respectively by multiplication, addition and complementation among the variables of the theory.
-This has an interesting physical significance, since the variables associated to magnitudes of physical systems can be added and multiplied together---i.e.\ one can sum and multiply measurement results.
+The idea is that the logical operations of conjunction, disjunction and negation correspond in the context of classical mechanics respectively to multiplication, addition and complementation among the observables of the theory.
+This has an interesting physical significance, since these observables associated to magnitudes of physical systems can be added and multiplied together---i.e.\ one can sum and multiply measurement results: ``since the observables [of CM] form a commutative Poisson algebra, addition and multiplication of observables reflect the action of adding and multiplying results of different measurements'' (cf.\ \cite{David:2015}, p.\ 77).
 Both propositional logic and classical mechanics, thus, generate a complemented Boolean lattice; from this fact it follows that the observable algebra of CM ``is isomorphic to a Boolean algebra of propositions with $(\wedge, \vee, \neg)$'' (\cite{David:2015}, p.
 79).
-In this regard, Beltrametti claims that 
+To this regard, Beltrametti claims that
 \begin{quote}
@@
 A point in phase space corresponds to the specification of the position and momentum variables of a certain classical system, whereas in QM the points of $\mathcal{H}$ correspond to wave functions.
-The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcomes in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.} as underlined by Svozi:
+The authors, then, find a connection between subsets of the observation-space of a system and subsets of its Hilbert space, specifying that quantum experimental propositions are mathematically represented by a closed linear subspace of $\mathcal{H}$; this step is crucial in order to obtain the quantum propositional calculus.\ Alternatively, we can say that quantum mechanical operators correspond to propositions with ``yes/no'' (``true/false'') outcomes in a logical system\footnote{Cf.\ also \cite{David:2015}, p.\ 78, where we read that ``An orthogonal projector $\textbf{P}$ onto a linear subspace $P\subset\mathcal{H}$ is indeed the operator associated to an observable that can take only the values 1 (and always 1 if the state $\psi\in P$ is in the subspace $P$) or 0 (and always 0 if the state $\psi\in P^{\perp}$ belongs to the orthogonal subspace to $P$).\ Thus we can consider that measuring the observable $\textbf{P}$ is equivalent to perform a test on the system, or to check the validity of a logical proposition $p$ on the system''.} as underlined by Svozil:
 \begin{quote}
@@
 \begin{quote}
-``The physical system has a property corresponding to the associated closed linear subspace'' (\cite{Svozi:1999}, p.\ 1).
+``The physical system has a property corresponding to the associated closed linear subspace'' (\cite{Svozil:1999}, p.\ 1).
 \end{quote}
@@
 Such a statement corresponds to the ``smallest closed subspace containing both $P$ and $Q$'' (\cite{Bacciagaluppi:2009}, p.\ 55).
-\item The logical implication is defined by set theoretic inclusion: given two experimental propositions $p$ and $q$, 
-$p$ implies $q$ means that whenever one predicts $p$ with certainty, one can predict also $q$ with certainty, and this is equivalent to stating that $p$ is a subset of $q$.
+\item As far as logical implication is concerned, \cite{vonNeumann:1936} claim that given two experimental propositions $p$ and $q$ about a certain physical system, ``$p$ implies $q$'' means that ``whenever one can predict $p$ with certainty, one can predict also $q$ with certainty'' (p.\ 827) and this is equivalent to stating that ``the mathematical representative of $p$ is a subset of the mathematical representative of $q$'' (\emph{ibid.}).
 This fact is particularly important since the authors showed that ``it is \emph{algebraically} reasonable to try to correlate physical qualities with subsets of phase-space'' and thus, ``physical qualities attributable to any physical system form a partially ordered system'' (\cite{vonNeumann:1936}, p.\ 828, notation adapted).

### 2021-10-26 Articles/2021-ReversePhysics/ReversePhysics.tex
ReversePhysics: draft with a few reviews

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{assumptionsofphysics}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\newcommand\partitle[1]{\textsc{#1}.}

\begin{document}

\title{Reverse Physics}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics. The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program. This allows us to identify physical starting points (as opposed to mathematical ones), compare their predictive power, and find those that can better serve as a foundation for our theories.
\end{abstract}

\maketitle

\section{Introduction}

%\begin{itemize}
%	\item Brief descriptions of the aims of foundations of mathematics
%	\item We want to follow a similar approach (not new interpretations of QM or new models for quantum gravity)
%	\item Reverse mathematics start from theorems and finds the smallest set of axioms needed to rederive them. We want to take physical theories and find the smalles set of physical assumptions that are equivalent to the theories themselves.
%	\item This paper is meant to showcase the approach and how it can lead to renewed physical insights into theories that may be thought as already fully understood and those that are not. The main point of the paper is not the results per se, but the novel of the approach proposed. We hope that other people will see the value and pursue similar lines of investigations.
%	
%\end{itemize}

Within the foundations of mathematics, the reverse mathematics program aims to start from known theorems and determine what axioms are required to prove them. This helps understand the relative logical strength of the theorems and get a better sense of their relationships. We propose a similar program for physics. We start from known physical theories or elements of physical theories, and we try to find premises that enable us to rederive them. The goal is to find premises that are physically more intuitive, so that ideally the arguments can be carried out with or without the math, and develop concepts that are of physical significance across theories.

While one does find efforts that analyze the premises of physical theories, for example in the context of foundations of quantum mechanics, statistical mechanics or the search for quantum gravity, their aim is typically different. Some are focused on a particular field (e.g. quantum theory) or try to demonstrate the centrality of one aspect (e.g. information) with regard to others. Some concentrate more on mathematical premises or structure, making the physical significance less than clear. Many aim to develop new theories, instead of first gaining further insight within current ones. We propose a more generalist approach that examines the relationships of physical ideas across subfields, which is best explained through examples.

To give a sense of the breadth of the reverse physics approach, we show how it can be used to pursue three different objectives. First, we illustrate multiple instances of reformulation. We study Hamilton's equation for a single degree of freedom to find six alternative characterizations that will connect different ideas from different fields under the same theme of determinism and reversibility. Next, we illustrate dependence analysis. We show that the uncertainty principle is not a consequence of the full quantum theory, but simply of its zero entropy bound of pure states. If a similar bound is imposed in classical mechanics, similar uncertainty relationships are recovered. Lastly, we illustrate reconceptualization. We take the third law of thermodynamics and rework it so that the role of crystalline substances to define zero entropy is instead played by the ``empty system''.

We want to stress that this paper is not about the originality of each single result, but about how these are used to solve a more general and higher goal: to uncover the core physical concepts and assumptions in the known physical theories and understand their generative power and their interrelationships.

%Use "investigation trumps rigor"
%Somewhere, note that "physics" entropy does not coincide with "information" entropy. Physics entropy is information about states as quantified by invariants.

\section{Determinism, reversibility and Hamiltonian mechanics}

%\begin{itemize}
%	\item Start with the equations of HM
%	\item Show Liouville's theorem and wonder what it means. Jacobian is 1. Show how Jacobian for 1 d.o.f. is really a Poisson bracket.
%	\item Connect to statistical mechanics: volume is number of states. HM conserves the number of states. States are mapped one to one on a continuum. (discuss problems of counting on a continuum)
%	\item Show that we can go backwards on a single degree of freedom
%	\item Wonder what to do on multiple d.o.f. 
%	\item Show symplectic form as the geometrical description. Generalize to d.o.f. by taking outer product. Show that areas remain orthogonal
%	\item What does orthogonality mean? Independence of d.o.f. (volume product of areas)
%	\item Show that preservation of the geometry (areas and perpendicularity) under time evolution gives us HM
%	\item HM is det/rev motion of indepent d.o.f.
%	\item We used count of states, log count of states is also conserved. What about information entropy? It is also conserved. 
%	\item Show we can proceed the other way. Det/rev means information before and after is the same. Information needs to be conserved -> Jacobian = 1 -> HM.
%	\item Note how bijection is not enough. There is an additional constraint. Why?
%	\item Measurements on a continuum have uncertainty. Connect to statistics and use the covariance matrix to describe a gaussian peak. The determinant of the matrix with the Jacobian. If jacobian is 1, determinant is the same, same precision of the description. HM <-> to requiring a bijection at the same level of precision.
%	\item Wonder why we are getting conservation of energy from det/rev, while we typically associate conservation of energy with system isolation (in the thermodynamic sense). Argue that if a system is deterministic and reversible, it means that the future/past state depends only on the system itself. The system does not depend on anything else -> it is isolated. There is a deep connection between isolation and det/rev
%	
%\end{itemize}

Let us start our discussion with the equations of motion given by classical Hamiltonian mechanics:
\begin{align}\label{HamiltonEquations1}
	\frac{dq}{dt} = \frac{\partial H}{\partial p} \; \; \; \; \; \; \frac{dp}{dt} = - \frac{\partial H}{\partial q}
\end{align}
Our first task is to find a set of mathematical conditions that are equivalent to these equations.

Let us group the state variables $\xi^a = \{q, p\}$ and consider the displacement vector field $S = \frac{d\xi^a}{dt} = \{ \frac{dq}{dt}, \frac{dp}{dt} \} $ which tells the direction in phase space in which each state moves over time. We find that $S$ is divergenceless:
\begin{align}\label{divergenceless}
\frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} =0
\end{align}
meaning that the flow through any closed region in phase space is zero. We also find the converse is true: a two dimensional divergenceless field $S$ always allows a stream function $H$ such that
\begin{align}
	S^q = \frac{\partial H}{\partial p} \; \; \; \; \; \; S^p = - \frac{\partial H}{\partial q}
\end{align}
In other words, Hamiltonian evolution for one degree of freedom is exactly the evolution for which the flow over a closed region is zero.

Let us now look at how an infinitesimal region of phase space evolves. Suppose $Q = q + \frac{dq}{dt} dt = q + S^q dt$ and $P = p + \frac{dp}{dt} dt = p + S^p dt$. The infinitesimal region of area $dq dp$ will become $dQ dP = |J| dq dp$ where:
\begin{align}
	|J| = \begin{vmatrix}
		\frac{\partial Q}{\partial q} & \frac{\partial Q}{\partial p} \\
		\frac{\partial P}{\partial q} & \frac{\partial P}{\partial  p} 
	\end{vmatrix} = \frac{\partial Q}{\partial q} \frac{\partial P}{\partial  p} - \frac{\partial P}{\partial q} \frac{\partial Q}{\partial p}
\end{align}
Note how the Jacobian determinant, in this simple case, coincides with the Poisson bracket. With a simple substitution we find:
\begin{align}
	|J|\approx 1 + \left( \frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} \right)dt.
\end{align}
Note that the first order term is exactly the divergence, therefore condition \eqref{divergenceless} is equivalent to
\begin{align}\label{UnitaryJacobian}
	|J| = 1
\end{align}
which means the initial area $dqdp$ is equal to the final area $dQdP$.

So we have found that, for a two dimensional manifold, Hamiltonian evolution \eqref{HamiltonEquations1}, zero flow over closed regions \eqref{divergenceless} and preservation of area \eqref{UnitaryJacobian} are mathematically the same condition.\footnote{This is essentially a short proof for Liouville's theorem that can work in both directions.} However, we still have to answer the main question: what does this mean physically?

To do that, we look at statistical mechanics. There we use the area of a phase space region to count the number of states. Area preservation, then, means that the evolution preserves the state count: we start and we end with the same number of states. We have a one-to-one map between past and future states. Given an initial state we have one and only one final state. Physically, this means
\begin{align}\label{DeterminismReversibility}
	\text{Deterministic and reversible evolution}
\end{align}
is equivalent to conditions \eqref{HamiltonEquations1}, \eqref{divergenceless} and \eqref{UnitaryJacobian}. But determinism and reversibility is a physical condition, a physical property of the evolution: we have a physical characterization of Hamiltonian evolution.

This seems to tell us that the physical requirement of deterministic and reversible evolution over the continuum mathematically is not simply a bijection. The reason is that, over a continuum, counting points is not enough. A segment a meter long has as many points as a segment a kilometer long, but they are not the same length. In the same way, all finite regions of phase space have infinitely many points, but physically they do not contain the same number of states. Mathematically, we need a measure to give a size to each region, and therefore deterministic and reversible evolution is a bijection that preserves the measure, that preserves the state count: Hamiltonian evolution.

Note that this makes sense from a statistical mechanics/thermodynamics perspective as well. In statistical mechanics, the entropy is the logarithm of the count of states, therefore conservation of number of states means conservation of entropy. This means a deterministic evolution that is thermodynamically reversible (conservation of entropy) is also reversible in the dynamical sense (conservation of state count). Therefore 
\begin{align}\label{DeterminismThermodynamicReversibility}
	\parbox{2.8in}{Deterministic and thermodynamically reversible evolution}
\end{align}
is yet another equivalent condition.

What about entropy in the information sense? If the evolution is deterministic and reversible, the amount of information needed to specify the initial state should be exactly the same as the amount of information needed to specify the final state. That is, if $I[\rho] = - \int \rho \log \rho dqdp$ is the Gibbs/Shannon entropy, we would expect the following condition:
\begin{align}\label{EntropyConservation}
	I[\rho(t_1)] = I[\rho(t_2)]
\end{align}
to be equivalent to the others. We find:
\begin{align}
	I[\rho(t + dt)] = I[\rho(t)] - \int \rho \log |J| dqdp.
\end{align}
Since the Jacobian determinant of a continuous transformation cannot be negative, condition \eqref{EntropyConservation} is indeed equivalent to \eqref{UnitaryJacobian} and therefore to all the others.

It may seem odd that conservation of entropy for a deterministic system gives us energy conservation: we typically associate energy conservation with system isolation in thermodynamics. Where is the connection? For a deterministic and reversible system, future and past states depend only on the state of the present system. Therefore they do not depend on the state of other systems or of the environment. This means that deterministic and reversible systems are necessarily isolated. This time, running the argument in the opposite direction presents a problem. One needs to first decide whether a non-deterministic isolated system makes sense. If one assumes that the source of non-determinism, of stochastic uncertainty, is always interaction with other systems, other degrees of freedom, than every isolated system is deterministic. In this case, we can run the argument in reverse: every isolated system is deterministic and reversible. Without that extra assumption, this reverse direction is not guaranteed.

Finally, let us go back to the idea that a bijection is not enough to characterize deterministic and reversible evolution. Preservation of a measure is a mathematical idea. Is there a more physical way to look at it? The issue is that scientific measurements over a continuum can only carry finite precision. When considering how the information goes back and forth in time, then, we also need to take into account how the finite precision is mapped. While it may be true that the evolution of a damped harmonic oscillator is a bijection, the points get denser and denser around the equilibrium. Once we are close to equilibrium it becomes impossible to tell when the oscillator was started: the damped harmonic oscillator is not reversible in any practical sense if precision is taken into consideration. Let us see if this argument can be run formally. We want a coordinate invariant quantity that characterizes uncertainty. The obvious choice is the determinant of the covariance matrix:
\begin{align}\label{CovarianceMatrix}
	|\Sigma| = \begin{vmatrix}
		\sigma_q^2 & cov_{q,p} \\
		cov_{p,q} & \sigma_p^2
	\end{vmatrix} = \sigma_q^2 \sigma_p^2 - cov_{q,p}^2
\end{align}
We can therefore imagine the following condition
\begin{align}\label{BijectionUncertainty}
	|\Sigma(t_1)| = |\Sigma(t_2)| 
\end{align}
for which the uncertainty remains constant in time. Is this yet another equivalent condition? If we assume the spread of the distribution is small enough, we find
\begin{align}
	|\Sigma(t + dt)| = |J| |\Sigma(t)| |J|.
\end{align}
therefore \eqref{BijectionUncertainty} is indeed equivalent to \eqref{UnitaryJacobian} and to all others.

We could go on to find other relationships and extend the ones we found to multiple degrees of freedom, but we believe this should be enough to highlight the power of the reverse physics approach. The first thing to note is how in a couple of pages we have found fundamental connections between classical Hamiltonian mechanics \eqref{HamiltonEquations1}, vector calculus \eqref{divergenceless}, differential geometry \eqref{UnitaryJacobian}, statistical mechanics \eqref{DeterminismReversibility}, thermodynamics \eqref{DeterminismThermodynamicReversibility}, information theory \eqref{EntropyConservation} and plain statistics \eqref{BijectionUncertainty}. This helps foster a sense of unity of these disparate disciplines and their perspectives, a sense that is sorely lacking both in research and education. We believe that a single unified view of this kind will bring more coherence and clarity to physics than, for example, a unified theory for the fundamental forces: nature is one, and does not care about how we have divided academic knowledge.

Another advantage is that we were able to carry out many of the arguments conceptually. Mathematics is then used to better articulate the physical arguments. This should please those who believe that physics, not mathematics, should be more at the center of the discussion.

The other interesting aspect is how much we were able to find in theories that are generally considered well understood. While it is common for some people to be aware of some of these results, we have found that most results are unknown to most people.

\section{Uncertainty principle revisited}

%\begin{itemize}
%	\item We saw that classical mechanics is more than just points: it says something about uncertainty as well. Quantum mechanics has an uncertainty principle. Is there any connections?
%	\item Since there was also a connection to entropy, turn to entropy in quantum mechanics. Entropy is bounded from below: all pure states have zero entropy. In classical mechanics, however, we can prepare distributions with arbitrarily small entropy. Can the fact that the entropy is bound from below in QM be the source of the uncertainty?
%	\item Assume entropy is bounded from below in CM. Show you get uncertainty relationship in a single degree of freedom.
%	\item Uncertainty relationship <-> bounded entropy = bounded information
%\end{itemize}

Let us now turn our attention to quantum mechanics, and see if our approach can shed new light on a theory that is not generally considered to be well understood. Since we talked about the role of uncertainty in classical mechanics, let us concentrate on the uncertainty principle which states that every state has to satisfy the relationship
\begin{align}\label{UncertaintyPrinciple}
	\sigma_q \sigma_p \geq \frac{\hbar}{2}.
\end{align}

We can start with an interesting observation. If we look back at \eqref{CovarianceMatrix}, we have:
\begin{align}
	\sigma_q^2 \sigma_p^2 = |\Sigma| + cov_{q,p}^2 \geq |\Sigma|.
\end{align}
Since by \eqref{BijectionUncertainty} $|\Sigma|$ is a constant of motion, we find that during Hamiltonian evolution the product of the uncertainty is bounded. The lowest uncertainty is reached when there are no correlations and the covariance is zero. Though it is a conceptually different relation, since $|\Sigma|$ is just a constant of motion, it gives us the following intuition: a coordinate invariant cap to the uncertainty will produce an inequality on the product of variances. Where can we find such a cap in quantum mechanics?

Let us turn our attention to entropy, represented by the von Neumann entropy $I[\rho] = - tr (\rho \log \rho)$. For every pure state $|\psi \rangle$ we have the following property:
\begin{align}\label{ZeroQuantumEntropy}
	\begin{split}
		I[|\psi\rangle\langle\psi|] = 0.
	\end{split}
\end{align}
In other words, all pure states have zero entropy, the entropy is capped. Since entropy and uncertainty are related, is fixing the amount of entropy enough to recover the uncertainty principle?

We can test that hypothesis by studying the space of classical distributions with a fixed value of entropy $I_0$. We find that they have to satisfy the relationship\footnote{The careful reader will note the units do not quite work. The issue here is the $\log \rho$ in the Shannon/Gibbs entropy, as $\rho$ is not a pure number. Introducing a dimensionful constant $\log \rho \hat{h}$ would fix the expression, which would then fix the uncertainty relationship as well. } (TODO)
\begin{align}\label{ClassicalUncertaintyPrinciple}
	\sigma_q \sigma_p \geq \frac{e^{I_0}}{2 \pi e}.
\end{align}
The equality holds if the distribution is the product of two independent Gaussians.

The following condition:
\begin{align}\label{BoundedEntropy}
	\text{The entropy of the system is fixed}
\end{align}
is therefore enough to recover an uncertainty relationship independently of the theory. This can be understood as fixing the amount of information carried by the state. The reverse argument does not quite work, so condition \eqref{BoundedEntropy} is a stronger condition than the existence of an uncertainty relationship.

This example shows another advantage of reverse physics: it allows us to pinpoint what part of a theory is responsible for which effect. In this case, we have found that it is not quantum mechanics per se that leads to uncertainty relationships, just the entropy cap on pure states. If we implemented a similar cap in classical mechanics, we would obtain the same result. Therefore condition \eqref{ZeroQuantumEntropy} is a more fundamental and clear way to characterize the uncertainty in quantum mechanics, making it evident that all pure states carry the same information, they describe the system at the same level of precision.

\section{A new understanding of the third law}

%\begin{itemize}
%	\item We saw that in quantum mechanics entropy is bounded from below. Why?
%	\item Turn to thermodynamics. The third also says that thermodynamic entropy (not information entropy) is bounded from below. The zero thermodynamic entropy, however, is formulated in terms of a pure crystal at zero temperature. This does not seem to be a fundamental characterization. Can we do better?
%	\item Note how entropy is an extensive quantity. Now consider the empty system. Since the composition of any system with the empty system must yield the same system, it also must yield the same entropy. -> Entropy of empty system must be zero. Better more fundamental characterization of zero entropy.
%	\item What would a system with negative entropy correspond to? To a description of a system that is more precise than saying the system is not there. But that is absurd! Therefore, it is a logical necessity that the entropy of a system is bound from below.	
%\end{itemize}

In the previous section we saw the lower entropy bound of zero entropy is built into quantum mechanics and not in classical mechanics. However, the bound is also built into thermodynamics through the third law\footnote{This formulation is given by Gilbert Newton Lewis and Merle Randall, Thermodynamics and the Free Energy of
	Chemical Substances. McGraw-Hill, New York, 1923}
\begin{align}\label{ThirdLaw}
	\parbox{2.8in}{Every substance has a finite positive entropy, but
		at the absolute zero of temperature the entropy may become zero, and does so
		become in the case of perfect crystalline substances.}
\end{align}
Therefore we have two fundamental physical theories that share this trait. Can we find a more general characterization for this?

The first step is to look for a system that can conceptually function as a zero for entropy that feels a bit less arbitrary than a perfect crystalline substance at zero temperature. Let us recall that a fundamental property for thermodynamic entropy is that it is an extensive quantity, it is additive for independent systems:
\begin{align}
	I_{AB} = I_{A} + I_{B}.
\end{align}
Is there a system that acts as a zero for system composition? This would be the empty system $\emptyset$: any system combined with the empty system $\emptyset$ will remain unchanged.\footnote{If we consider systems as a monoid under composition, the empty system is the identity element, much like the number zero, the empty set or the identity map in their respective structures.} In particular, we must have:
\begin{align}\label{EntropyEmptySystem}
	I_{\emptyset} = I_{\emptyset\emptyset} = I_{\emptyset} + I_{\emptyset} = 0.
\end{align}
We can therefore use the empty system as our zero reference for entropy.

This move is a conceptual shift. In thermodynamics, the idea that the entropy for crystalline substances is zero is a phenomenological assumption: it is so because it justifies the behavior of substances as they approach absolute zero. We need statistical mechanics to explain it. Condition $\eqref{EntropyEmptySystem}$, however, is not a phenomenological consideration but a conceptual necessity: it must be so. In this sense, the new zero entropy system is more fundamental. But does it relate to the old ideas? In terms of statistical mechanics, the crystalline substance has zero entropy because it has one possible way to be. The empty system also has one possible configuration. So the old ideas indeed carry over: a crystalline substance will have to have the same entropy as the empty system.

Now that we have a better concept for zero entropy, we need to understand why we can't have states with lower entropy. Let us consider it from an information theoretical perspective. Suppose we had a state with negative entropy. This would mean that it is better specified than a state with zero entropy. The state is better specified than saying that the system is not there. But stating that the system is not there already completely specifies the system. Therefore the condition that imposes a lower bound of zero for entropy is the following:
\begin{align}\label{NewThirdLaw}
	\parbox{2.8in}{No state can describe a system more accurately than stating the system is not there in the first place.}
\end{align}
Again, note that this condition is not phenomenological: it is a logical necessity. This condition is realized in quantum mechanics: pure states represent the most precise descriptions of the system and they have zero entropy much like the vacuum, another pure state. This condition is realized in thermodynamics through the third law. In fact, the third law could be rederived as follows. Assuming \eqref{NewThirdLaw} tells us that entropy is bounded at zero. Since the entropy is a convex function of energy, entropy will reach its minimum at zero temperature, which is the only temperature that can reach zero entropy. We then use the previous result that the entropy of a crystalline substance at zero temperature is the same of an empty system.

This showcases the final advantage of reverse physics. We want to differentiate between ``assumptions'' (conditions that are not necessarily true) and ``principles'' (those that have to be taken to be true); we want to differentiate between conditions that are ``phenomenological'' (empirically derived or justified) from those that are ``conceptual'' (logical necessities or consequences of definitions).\footnote{Philosophers may call these ``constitutive'' conditions.} If we are able to elevate assumptions to principles, or replace phenomenological assumptions with conceptual ones, then the premises for our theories are sturdier and we have improved their foundations.

%\section{A more unified physics}
%
%\begin{itemize}
%	\item Put physics back at the center of the discussion (we are exploring physical concepts and their relationships, the math is there to carry out the arguments)
%	\item Appeal to tradition (thinking about small set of physical assumptions was what foundations of classical mechanics, thermodynamics, ... was about)
%	\item Conceptual unification vs theoretical unification
%	\item Common standard tools (like set theory, category theory, ...)
%	\item Connection between fields
%	\item Reduce the number of fundamental concepts
%	\item These more powerful tools may help theoretical unification
%	\item realm of validity of the different theoretical theories
%	\item realm of validity of the underlying mathematical tools
%\end{itemize}

\section{Conclusion}

We have shown how the reverse physics approach, trying to find different premises from which to rederive physical theories or specific results, can help us improve our understanding. In the first example we showed how we can take a single theory (classical Hamiltonian mechanics) and characterize it in several ways consistent with each other (deterministic evolution that is either dynamically or thermodynamically reversible, conservation of information entropy, conservation of measurement precision, ...). We have seen how this helps foster a unified understanding of physics, and puts physics (and not mathematics) at the center of the discussion. We then showed how we can isolate a specific result of a theory (the uncertainty relationship), see what specific part of the theory is responsible for that result (fixing the entropy) and showed how another theory (classical mechanics) can be modified to achieve the same result. Lastly, we saw how we can recast a mostly phenomenological condition (the zero entropy of a crystalline structure at zero temperature) into a conceptual one (the zero entropy of an empty system).

We believe this type of work can be beneficial in several ways. At the very least, it can have a significant impact within physics education. Most importantly, it leads to common ideas and definitions that span different fields of physics, which may help develop common tools, much like set theory and category theory provide standard tools for all of mathematics. Lastly, as it helps us understand the realm of validity of the different theories and also the underlying mathematical tools, it may give insight into the development of new tools and new theories.

More than the particular results presented in this paper, it is the type of approach that we want to promote. We hope others will present their own reverse physics results, even better than those presented here, and that this may lead to a renewed interest in the manner in which physicists in the past tried to formulate the foundations of our field: through physical principles, laws and assumptions.

\section{Proofs and calculations}

We omitted all proofs from the main body of the text because, as we stated, the physics should be given full attention. However, we do have to show that our arguments are on solid ground, so we include here basic proofs for all our major claims. To keep the discussion light, we will use the standards of rigor typically employed in physics. Note that most of the derivations are trivial, some already included in many textbooks. However, given the interdisciplinary nature of our subject, we find it useful for the reader to collect them all here.

\begin{prop}
	A dynamical system characterized by two variables $\{q, p\}$ follows Hamilton's equations if and only if the displacement field $S=\{\frac{dq}{dt}, \frac{dq}{dt} \}$ is divergenceless.
\end{prop}

\begin{proof}
	Suppose the evolution is Hamiltonian. Then $S=\{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$. The divergence is $div(S) = \frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} = \frac{\partial \partial H}{\partial q\partial p} - \frac{\partial \partial H}{\partial p\partial q} = 0$.

	Conversely, suppose $S$ is divergenceless. Then, it admits a stream function $H$ such that $\{\frac{dq}{dt}, \frac{dq}{dt} = \{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$, which are Hamilton's equations
\end{proof}

\begin{prop}
	The displacement field $S=\{\frac{dq}{dt}, \frac{dq}{dt} \}$ of a dynamical system characterized by two variables $\{q, p\}$ is divergenceless if and only if the Jacobian determinant of the infinitesimal time evolution is equal to one.
\end{prop}

\begin{proof}
	Let $Q = q + \frac{dq}{dt} dt = q + S^q dt$ and $P = p + \frac{dp}{dt} dt = p + S^p dt$. The Jacobian determinant is given by:
	\begin{align*}
		|J| &= \begin{vmatrix}
			\frac{\partial Q}{\partial q} & \frac{\partial Q}{\partial p} \\
			\frac{\partial P}{\partial q} & \frac{\partial P}{\partial  p} 
		\end{vmatrix} = \frac{\partial Q}{\partial q} \frac{\partial P}{\partial  p} - \frac{\partial P}{\partial q} \frac{\partial Q}{\partial p} \\
		&= \left(1 + \frac{\partial S^q}{\partial  q}dt\right)\left(1 + \frac{\partial S^p}{\partial  p}dt\right) - \left(\frac{\partial S^p}{\partial  q}dt\right)\left(\frac{\partial S^q}{\partial  p}dt\right) \\
		&= 1 + \left(\frac{\partial S^q}{\partial  q} + \frac{\partial S^p}{\partial  p}\right)dt + O(dt^2)
	\end{align*}
	The Jacobian determinant will be 1 if and only if the first order term is zero. Since the first order term is the divergence of $S$, this proves the proposition.	
\end{proof}

\begin{remark}
	An unstated assumption here is that $H$ is twice differentiable. The Jacobian and the divergence, in fact, would not be well defined if $H$ were differentiable only. In that case, the acceleration $a = \frac{d}{dt}\frac{dq}{dt} = \frac{d}{dt} \frac{\partial H}{\partial p}$ would also be ill defined. With this in mind, if a Hamiltonian is not twice differentiable at a point it is better to consider the system not Hamiltonian at that point.
\end{remark}

\begin{prop}
	The information entropy is conserved during a continuous transformation if and only if the Jacobian is unitary.
\end{prop}

\begin{proof}
	Let $I[\rho(x^i)] = - \int \rho(x^i) \log \rho(x^i) \, dx^n$ be the Shannon entropy of the distribution $\rho(x^i)$. Let $y^j=y^j(x^i)$ be a differentiable transformation. Since $\rho(x^i)$ is a density, we have $\rho(y^j) = \rho(x^i) |J|$. The information entropy after the transformation will be
	\begin{align*}
		I[\rho(y^j)] &= - \int \rho(y^j) \log \rho(y^j) \, dy^n \\
		&= - \int \rho(x^i) |J| \log \left( \rho(x^i) |J| \right) \, dy^n \\
		&= - \int \rho(x^i) \log \left( \rho(x^i) |J| \right) \, dx^n \\
		&= - \int \rho(x^i) \log \rho(x^i) \, dx^n - \int \rho(x^i) \log |J| \, dx^n \\
		&= I[\rho(x^i)] - \int \rho(x^i) \log |J| \, dx^n.
	\end{align*}
	The information entropy to be conserved for every $\rho$ exactly when $|J|$ is $\pm 1$ everywhere. Given that the transformation is continuous, the Jacobian determinant must be positive (i.e. a continuous transformation cannot be a reflection) and therefore the entropy is conserved if and only if the Jacobian is unitary.
\end{proof}

\begin{prop}
	A continuous transformation will conserve the uncertainty of a distribution with small support if and only if the Jacobian is unitary. The uncertainty is characterized by the determinant of the covariance matrix and small means the evolution can be considered approximately linear over the support of the distribution.
\end{prop}

\begin{proof}
	As we assume the distribution to be small enough, we can linearize the transformation to $y^j = J^j_i x^i + B^j$ where $J$ is the Jacobian. Noting that the covariance is a linear operator that does not depend on the expectation of the variables, we have
	\begin{align*}
		|\Sigma^{ij}| &= |cov(y^i, y^j)| \\
		&= |cov(J^i_k x^k + B^i, J^j_l x^l + B^j)| \\
		&= |cov(J^i_k x^k , J^j_l x^l )| \\
		&= |J^i_k cov(x^k , x^l ) J^j_l| \\
		&= |J^i_k | | \Sigma^{kl} | | J^j_l| \\
	\end{align*}
	The uncertainty is conserved for all small distributions exactly when $|J|$ is $\pm 1$ everywhere. Given that the transformation is continuous, the Jacobian determinant must be positive (i.e. a continuous transformation cannot be a reflection) and therefore the entropy is conserved if and only if the Jacobian is unitary.
\end{proof}

\begin{prop}
	The von Neumann entropy of a pure state is zero.
\end{prop}

\begin{proof}
	Let $\rho$ be the density matrix of a pure state. We have $\rho = |\psi \rangle \langle psi | = |\psi \rangle \langle psi |\psi \rangle \langle psi | = \rho^2$. The von Neumann entropy is $I[\rho] = - tr (\rho \log \rho) = - tr (\rho \log \rho^2) = - 2 tr (\rho \log \rho) = 2 I[\rho]$. Therefore $I[\rho] = 0$.
\end{proof}

\begin{prop}
	Let $\rho(q,p)$ be a normalized density distribution over the 2-dimensional manifold charted by $(q,p)$. Furthermore, let $I_0$ be the value of the Shannon/Gibbs entropy $I[\rho]$. Then
	$$ 	\sigma_q \sigma_p \geq \frac{e^{I_0}}{2 \pi e}.$$
	Furthermore, the equal sign applies in the case where $\rho$ is the product of two gaussians.
\end{prop}

\begin{proof}

\end{proof}

\bibliography{bibliography}

\end{document}

### 2021-11-08 Articles/2021-ReversePhysics/ReversePhysics.tex
ReversePhysics: review before arxiv

@@
 The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program.
-This allows us to identify physical starting points (as opposed to mathematical ones), compare their predictive power, and find those that can better serve as a foundation for our theories.
+This allows us to identify physical starting points (as opposed to mathematical ones), compare their deductive power, and find those that can better serve as a foundation for our theories.
 \end{abstract}
@@
 
-Within the foundations of mathematics, the reverse mathematics program aims to start from known theorems and determine what axioms are required to prove them.
+Within the foundations of mathematics, the reverse mathematics program\cite{friedman1976systems,simpson2017reverse,stillwellreverse} aims to start from known theorems and determine what axioms are required to prove them.
 This helps understand the relative logical strength of the theorems and get a better sense of their relationships.
@@
 
-While one does find efforts that analyze the premises of physical theories, for example in the context of foundations of quantum mechanics, statistical mechanics or the search for quantum gravity, their aim is typically different.
+While one does find efforts that analyze the premises of physical theories (see e.g.
+Refs.~\cite{chiribella2011informational, selby2021reconstructing, giles2016mathematical, boyling1972axiomatic, PhysRev.101.860, Haag1964848, axiomaticQFT1975, masanes2019measurement, carcassi2021four}), for example in the context of foundations of quantum mechanics, statistical mechanics or the search for quantum gravity, their aim is typically different.
 Some are focused on a particular field (e.g.
@@
 To give a sense of the breadth of the reverse physics approach, we show how it can be used to pursue three different objectives.
-First, we illustrate multiple instances of reformulation.
-We study Hamilton's equation for a single degree of freedom to find six alternative characterizations that will connect different ideas from different fields under the same theme of determinism and reversibility.
-Next, we illustrate dependence analysis.
-We show that the uncertainty principle is not a consequence of the full quantum theory, but simply of its zero entropy bound of pure states.
+First, we illustrate multiple instances of \emph{reformulation}.
+We study Hamilton's equations for a single degree of freedom to find six alternative characterizations that will connect different ideas from different fields under the same theme of determinism and reversibility.
+Next, we illustrate \emph{dependence analysis}.
+We show that the uncertainty principle is not a consequence of the full quantum theory, but simply of its zero entropy bound on pure states.
 If a similar bound is imposed in classical mechanics, similar uncertainty relationships are recovered.
-Lastly, we illustrate reconceptualization.
+Lastly, we illustrate \emph{reconceptualization}.
 We take the third law of thermodynamics and rework it so that the role of crystalline substances to define zero entropy is instead played by the ``empty system''.
 
-We want to stress that this paper is not about the originality of each single result, but about how these are used to solve a more general and higher goal: to uncover the core physical concepts and assumptions in the known physical theories and understand their generative power and their interrelationships.
+As we want the discussion to be focused on the physics, we will use the most widely known math among physicists and relegate mathematical derivations to the appendix.
+We also want to stress that this paper is not about the originality of each single result, but about how these are used to address a more general and higher goal: to uncover the core physical concepts and assumptions in the known physical theories and understand their generative power and their interrelationships.
 
@@
 \end{align}
-In other words, Hamiltonian evolution for one degree of freedom is exactly the evolution for which the flow over a closed region is zero.
+In other words, Hamiltonian evolution for one degree of freedom is exactly the evolution for which the flow over a closed region in phase space is zero.
 
@@
 One needs to first decide whether a non-deterministic isolated system makes sense.
-If one assumes that the source of non-determinism, of stochastic uncertainty, is always interaction with other systems, other degrees of freedom, than every isolated system is deterministic.
+If one assumes that the source of non-determinism, of stochastic uncertainty, is always interaction with other systems, other degrees of freedom, then every isolated system is deterministic.
 In this case, we can run the argument in reverse: every isolated system is deterministic and reversible.
@@
 The first thing to note is how in a couple of pages we have found fundamental connections between classical Hamiltonian mechanics \eqref{HamiltonEquations1}, vector calculus \eqref{divergenceless}, differential geometry \eqref{UnitaryJacobian}, statistical mechanics \eqref{DeterminismReversibility}, thermodynamics \eqref{DeterminismThermodynamicReversibility}, information theory \eqref{EntropyConservation} and plain statistics \eqref{BijectionUncertainty}.
-This helps foster a sense of unity of these disparate disciplines and their perspectives, a sense that is sorely lacking both in research and education.
-We believe that a single unified view of this kind will bring more coherence and clarity to physics than, for example, a unified theory for the fundamental forces: nature is one, and does not care about how we have divided academic knowledge.
+This helps foster a sense of unity of these disparate disciplines and their perspectives, a sense that is sorely lacking both in research and education: nature is one, and does not care about how we have divided academic knowledge.
+We believe that a single unified view of this kind will bring more coherence and clarity to physics than, for example, a unified theory for the fundamental forces.
 
@@
 Mathematics is then used to better articulate the physical arguments.
-This should please those who believe that physics, not mathematics, should be more at the center of the discussion.
+This should please those who believe that physics, not mathematics, should be more at the center of the discussion.\cite{hossenfelder2018lost, woit2006not}
 
@@
 Let us now turn our attention to quantum mechanics, and see if our approach can shed new light on a theory that is not generally considered to be well understood.
-Since we talked about the role of uncertainty in classical mechanics, let us concentrate on the uncertainty principle which states that every state has to satisfy the relationship
+Since we talked about the role of uncertainty in classical mechanics, let us concentrate on the uncertainty principle, which states that every state has to satisfy the relationship
 \begin{align}\label{UncertaintyPrinciple}
@@
 \end{align}
-Since by \eqref{BijectionUncertainty} $|\Sigma|$ is a constant of motion, we find that during Hamiltonian evolution the product of the uncertainty is bounded.
+Since by \eqref{BijectionUncertainty} $|\Sigma|$ is a constant of motion, we find that during Hamiltonian evolution the uncertainty is bounded.
 The lowest uncertainty is reached when there are no correlations and the covariance is zero.
@@
 Introducing a dimensionful constant $\log \rho \hat{h}$ would fix the expression, which would then fix the uncertainty relationship as well.
-} (TODO)
+}
 \begin{align}\label{ClassicalUncertaintyPrinciple}
@@
 This can be understood as fixing the amount of information carried by the state.
-The reverse argument does not quite work, so condition \eqref{BoundedEntropy} is a stronger condition than the existence of an uncertainty relationship.
+The reverse argument does not work, so condition \eqref{BoundedEntropy} is a stronger condition than the existence of an uncertainty relationship.
 
@@
 In the previous section we saw the lower entropy bound of zero entropy is built into quantum mechanics and not in classical mechanics.
-However, the bound is also built into thermodynamics through the third law\footnote{This formulation is given by Gilbert Newton Lewis and Merle Randall, Thermodynamics and the Free Energy of
-	Chemical Substances. McGraw-Hill, New York, 1923}
+However, the bound is also built into thermodynamics through the third law\footnote{This formulation is given by Ref.
+\cite{lewis1923thermodynamics}}
 \begin{align}\label{ThirdLaw}
@@
 Assuming \eqref{NewThirdLaw} tells us that entropy is bounded at zero.
-Since the entropy is a convex function of energy, entropy will reach its minimum at zero temperature, which is the only temperature that can reach zero entropy.
-We then use the previous result that the entropy of a crystalline substance at zero temperature is the same of an empty system.
+Since the entropy in thermodynamics is a convex function of energy, entropy will reach its minimum at zero temperature, which is the only temperature that can reach zero entropy.
+We then use the previous result that the entropy of a crystalline substance at zero temperature is the same as for an empty system.
 
@@
 More than the particular results presented in this paper, it is the type of approach that we want to promote.
-We hope others will present their own reverse physics results, even better than those presented here, and that this may lead to a renewed interest in the manner in which physicists in the past tried to formulate the foundations of our field: through physical principles, laws and assumptions.
+We hope others will present their own reverse physics results, and that this may lead to a renewed interest in the manner in which physicists in the past tried to formulate the foundations of our field: through physical principles, laws and assumptions.
 
-\section{Proofs and calculations}
+\section*{Appendix: Proofs and calculations}
 
-We omitted all proofs from the main body of the text because, as we stated, the physics should be given full attention.
-However, we do have to show that our arguments are on solid ground, so we include here basic proofs for all our major claims.
-To keep the discussion light, we will use the standards of rigor typically employed in physics.
-Note that most of the derivations are trivial, some already included in many textbooks.
-However, given the interdisciplinary nature of our subject, we find it useful for the reader to collect them all here.
+We include here all proofs which we omitted from the main body since, as we stated, the physics should be given full attention.
+To keep the discussion light and accessible to the widest audience, we use the simplest mathematical techniques to get the result.
+Some readers may feel that some results are well known, or too simple to even be included.
+Due to the interdisciplinary nature of the subject, we believe it is useful to collect them all here.
 
 \begin{prop}
-	A dynamical system characterized by two variables $\{q, p\}$ follows Hamilton's equations if and only if the displacement field $S=\{\frac{dq}{dt}, \frac{dq}{dt} \}$ is divergenceless.
+	A dynamical system characterized by two variables $\{q, p\}$ follows Hamilton's equations if and only if the displacement field $S=\{\frac{dq}{dt}, \frac{dp}{dt} \}$ is divergenceless.
 \end{prop}
@@
 Then $S=\{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$.
-The divergence is $div(S) = \frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} = \frac{\partial \partial H}{\partial q\partial p} - \frac{\partial \partial H}{\partial p\partial q} = 0$.
+The divergence is $div(S) = \frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} = \frac{\partial^2 H}{\partial q\partial p} - \frac{\partial^2H}{\partial p\partial q} = 0$.
 
 	Conversely, suppose $S$ is divergenceless.
-Then, it admits a stream function $H$ such that $\{\frac{dq}{dt}, \frac{dq}{dt} = \{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$, which are Hamilton's equations
+Then, it admits a stream function $H$ such that $\{\frac{dq}{dt}, \frac{dp}{dt}\} = \{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$, which are Hamilton's equations.
 \end{proof}
@@
 \begin{prop}
-	The displacement field $S=\{\frac{dq}{dt}, \frac{dq}{dt} \}$ of a dynamical system characterized by two variables $\{q, p\}$ is divergenceless if and only if the Jacobian determinant of the infinitesimal time evolution is equal to one.
+	The displacement field $S=\{\frac{dq}{dt}, \frac{dp}{dt} \}$ of a dynamical system characterized by two variables $\{q, p\}$ is divergenceless if and only if the Jacobian determinant of the infinitesimal time evolution is equal to one.
 \end{prop}
@@
 	An unstated assumption here is that $H$ is twice differentiable.
-The Jacobian and the divergence, in fact, would not be well defined if $H$ were differentiable only.
+The Jacobian and the divergence, in fact, would not be well defined if $H$ were differentiable only once.
 In that case, the acceleration $a = \frac{d}{dt}\frac{dq}{dt} = \frac{d}{dt} \frac{\partial H}{\partial p}$ would also be ill defined.
@@
 	\end{align*}
-	The information entropy to be conserved for every $\rho$ exactly when $|J|$ is $\pm 1$ everywhere.
-Given that the transformation is continuous, the Jacobian determinant must be positive (i.e.
-a continuous transformation cannot be a reflection) and therefore the entropy is conserved if and only if the Jacobian is unitary.
+	The information entropy is conserved for every $\rho$ if and only if $|J|$ is 1 everywhere. Therefore the entropy is conserved if and only if the Jacobian is unitary.
 \end{proof}
@@
 	\end{align*}
-	The uncertainty is conserved for all small distributions exactly when $|J|$ is $\pm 1$ everywhere.
+	The uncertainty is conserved for all small distributions if and only if $|J|$ is $\pm 1$ everywhere.
 Given that the transformation is continuous, the Jacobian determinant must be positive (i.e.
@@
 	Let $\rho$ be the density matrix of a pure state.
-We have $\rho = |\psi \rangle \langle psi | = |\psi \rangle \langle psi |\psi \rangle \langle psi | = \rho^2$.
+We have $\rho = |\psi \rangle \langle \psi | = |\psi \rangle \langle \psi |\psi \rangle \langle \psi | = \rho^2$.
 The von Neumann entropy is $I[\rho] = - tr (\rho \log \rho) = - tr (\rho \log \rho^2) = - 2 tr (\rho \log \rho) = 2 I[\rho]$.
@@
 \begin{proof}
+	We set up a minimization problem using Lagrange multipliers.
+We want to minimize the product of the variance $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp$ while keeping $\rho$ normalized and fixing the entropy to $I_0$.
+We have:
+	\begin{align*}
+		L = &\int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp \\
+		&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) \\ &+ \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\
+		\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
+		\lambda_2 \ln \rho_{\mathcal{c}} = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
+		\rho_{\mathcal{c}} = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}
+	\end{align*}
+	We solve the multipliers and have:
+	\begin{align*}
+		\rho_{\mathcal{c}} = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
+		I_0 = &\ln (2\pi e\sigma_q\sigma_p) \\
+		\sigma_q \sigma_p &= \frac{e^{I_0}}{2 \pi e}
+	\end{align*}
+	This shows that the gaussian minimizes the spread at fixed entropy, therefore all other distributions must have a larger or equal spread. 
+\end{proof}
 
-\end{proof}
+\begin{remark}
+	Note that the inverse argument does not work: if we have a bound on the uncertainty, we cannot say anything about the entropy.
+All distributions with higher entropy will satisfy a higher bound, therefore the entropy is arbitrarily high.
+Moreover, we can find distributions with low entropy but with arbitrarily high spread.
+Therefore a bound on the uncertainty still allows any value for entropy.
+\end{remark}
 

### 2021-11-15 Articles/2021-ReversePhysics/Submissions/Nature/ReversePhysics.tex
ReversePhysics: preparing for Nature submission

--- 2021-11-08 Articles/2021-ReversePhysics/ReversePhysics.tex
@@
 \title{Reverse Physics}
-\author{Gabriele Carcassi, Christine A. Aidala}
+\author{Gabriele Carcassi, Christine A. Aidala}}
 \affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
@@
 
-As we want the discussion to be focused on the physics, we will use the most widely known math among physicists and relegate mathematical derivations to the appendix.
+As we want the discussion to be focused on the physics, we will use the most widely known math among physicists and relegate mathematical derivations to the Supplementary Information.
 We also want to stress that this paper is not about the originality of each single result, but about how these are used to address a more general and higher goal: to uncover the core physical concepts and assumptions in the known physical theories and understand their generative power and their interrelationships.
@@
 
-\section*{Appendix: Proofs and calculations}
-
-We include here all proofs which we omitted from the main body since, as we stated, the physics should be given full attention.
-To keep the discussion light and accessible to the widest audience, we use the simplest mathematical techniques to get the result.
-Some readers may feel that some results are well known, or too simple to even be included.
-Due to the interdisciplinary nature of the subject, we believe it is useful to collect them all here.
-
-\begin{prop}
-	A dynamical system characterized by two variables $\{q, p\}$ follows Hamilton's equations if and only if the displacement field $S=\{\frac{dq}{dt}, \frac{dp}{dt} \}$ is divergenceless.
-\end{prop}
-
-\begin{proof}
-	Suppose the evolution is Hamiltonian.
-Then $S=\{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$.
-The divergence is $div(S) = \frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} = \frac{\partial^2 H}{\partial q\partial p} - \frac{\partial^2H}{\partial p\partial q} = 0$.
-
-	Conversely, suppose $S$ is divergenceless.
-Then, it admits a stream function $H$ such that $\{\frac{dq}{dt}, \frac{dp}{dt}\} = \{\frac{\partial H}{\partial p}, - \frac{\partial H}{\partial q} \}$, which are Hamilton's equations.
-\end{proof}
-
-\begin{prop}
-	The displacement field $S=\{\frac{dq}{dt}, \frac{dp}{dt} \}$ of a dynamical system characterized by two variables $\{q, p\}$ is divergenceless if and only if the Jacobian determinant of the infinitesimal time evolution is equal to one.
-\end{prop}
-
-\begin{proof}
-	Let $Q = q + \frac{dq}{dt} dt = q + S^q dt$ and $P = p + \frac{dp}{dt} dt = p + S^p dt$. The Jacobian determinant is given by:
-	\begin{align*}
-		|J| &= \begin{vmatrix}
-			\frac{\partial Q}{\partial q} & \frac{\partial Q}{\partial p} \\
-			\frac{\partial P}{\partial q} & \frac{\partial P}{\partial  p} 
-		\end{vmatrix} = \frac{\partial Q}{\partial q} \frac{\partial P}{\partial  p} - \frac{\partial P}{\partial q} \frac{\partial Q}{\partial p} \\
-		&= \left(1 + \frac{\partial S^q}{\partial  q}dt\right)\left(1 + \frac{\partial S^p}{\partial  p}dt\right) - \left(\frac{\partial S^p}{\partial  q}dt\right)\left(\frac{\partial S^q}{\partial  p}dt\right) \\
-		&= 1 + \left(\frac{\partial S^q}{\partial  q} + \frac{\partial S^p}{\partial  p}\right)dt + O(dt^2)
-	\end{align*}
-	The Jacobian determinant will be 1 if and only if the first order term is zero. Since the first order term is the divergence of $S$, this proves the proposition.	
-\end{proof}
-
-\begin{remark}
-	An unstated assumption here is that $H$ is twice differentiable.
-The Jacobian and the divergence, in fact, would not be well defined if $H$ were differentiable only once.
-In that case, the acceleration $a = \frac{d}{dt}\frac{dq}{dt} = \frac{d}{dt} \frac{\partial H}{\partial p}$ would also be ill defined.
-With this in mind, if a Hamiltonian is not twice differentiable at a point it is better to consider the system not Hamiltonian at that point.
-\end{remark}
-
-\begin{prop}
-	The information entropy is conserved during a continuous transformation if and only if the Jacobian is unitary.
-\end{prop}
-
-\begin{proof}
-	Let $I[\rho(x^i)] = - \int \rho(x^i) \log \rho(x^i) \, dx^n$ be the Shannon entropy of the distribution $\rho(x^i)$.
-Let $y^j=y^j(x^i)$ be a differentiable transformation.
-Since $\rho(x^i)$ is a density, we have $\rho(y^j) = \rho(x^i) |J|$.
-The information entropy after the transformation will be
-	\begin{align*}
-		I[\rho(y^j)] &= - \int \rho(y^j) \log \rho(y^j) \, dy^n \\
-		&= - \int \rho(x^i) |J| \log \left( \rho(x^i) |J| \right) \, dy^n \\
-		&= - \int \rho(x^i) \log \left( \rho(x^i) |J| \right) \, dx^n \\
-		&= - \int \rho(x^i) \log \rho(x^i) \, dx^n - \int \rho(x^i) \log |J| \, dx^n \\
-		&= I[\rho(x^i)] - \int \rho(x^i) \log |J| \, dx^n.
-	\end{align*}
-	The information entropy is conserved for every $\rho$ if and only if $|J|$ is 1 everywhere. Therefore the entropy is conserved if and only if the Jacobian is unitary.
-\end{proof}
-
-\begin{prop}
-	A continuous transformation will conserve the uncertainty of a distribution with small support if and only if the Jacobian is unitary.
-The uncertainty is characterized by the determinant of the covariance matrix and small means the evolution can be considered approximately linear over the support of the distribution.
-\end{prop}
-
-\begin{proof}
-	As we assume the distribution to be small enough, we can linearize the transformation to $y^j = J^j_i x^i + B^j$ where $J$ is the Jacobian.
-Noting that the covariance is a linear operator that does not depend on the expectation of the variables, we have
-	\begin{align*}
-		|\Sigma^{ij}| &= |cov(y^i, y^j)| \\
-		&= |cov(J^i_k x^k + B^i, J^j_l x^l + B^j)| \\
-		&= |cov(J^i_k x^k , J^j_l x^l )| \\
-		&= |J^i_k cov(x^k , x^l ) J^j_l| \\
-		&= |J^i_k | | \Sigma^{kl} | | J^j_l| \\
-	\end{align*}
-	The uncertainty is conserved for all small distributions if and only if $|J|$ is $\pm 1$ everywhere.
-Given that the transformation is continuous, the Jacobian determinant must be positive (i.e.
-a continuous transformation cannot be a reflection) and therefore the entropy is conserved if and only if the Jacobian is unitary.
-\end{proof}
-
-\begin{prop}
-	The von Neumann entropy of a pure state is zero.
-\end{prop}
-
-\begin{proof}
-	Let $\rho$ be the density matrix of a pure state.
-We have $\rho = |\psi \rangle \langle \psi | = |\psi \rangle \langle \psi |\psi \rangle \langle \psi | = \rho^2$.
-The von Neumann entropy is $I[\rho] = - tr (\rho \log \rho) = - tr (\rho \log \rho^2) = - 2 tr (\rho \log \rho) = 2 I[\rho]$.
-Therefore $I[\rho] = 0$.
-\end{proof}
-
-\begin{prop}
-	Let $\rho(q,p)$ be a normalized density distribution over the 2-dimensional manifold charted by $(q,p)$. Furthermore, let $I_0$ be the value of the Shannon/Gibbs entropy $I[\rho]$. Then
-	$$ 	\sigma_q \sigma_p \geq \frac{e^{I_0}}{2 \pi e}.$$
-	Furthermore, the equal sign applies in the case where $\rho$ is the product of two gaussians.
-\end{prop}
-
-\begin{proof}
-	We set up a minimization problem using Lagrange multipliers.
-We want to minimize the product of the variance $\sigma_q^2 \sigma_p^2 \equiv \int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp$ while keeping $\rho$ normalized and fixing the entropy to $I_0$.
-We have:
-	\begin{align*}
-		L = &\int (q-\mu_q)^2 \rho_{\mathcal{c}} \, dqdp \int (p-\mu_p)^2 \rho_{\mathcal{c}} \, dqdp \\
-		&+ \lambda_1(\int \rho_{\mathcal{c}} dqdp - 1) \\ &+ \lambda_2(- \int \rho_{\mathcal{c}} \ln \rho_{\mathcal{c}} \, dqdp - I_0)\\
-		\delta L = &\int \delta \rho_{\mathcal{c}} [(q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 + \\ &\lambda_1 - \lambda_2 \ln \rho_{\mathcal{c}} - \lambda_2 ] dqdp = 0 \\
-		\lambda_2 \ln \rho_{\mathcal{c}} = &\lambda_1 - \lambda_2 + (q-\mu_q)^2 \sigma_p^2 + \sigma_q^2 (p-\mu_p)^2 \\
-		\rho_{\mathcal{c}} = &e^{\frac{\lambda_1 - \lambda_2}{\lambda_2}}e^{\frac{(q-\mu_q)^2 \sigma_p^2}{\lambda_2}}e^{\frac{\sigma_q^2 (p-\mu_p)^2}{\lambda_2}}
-	\end{align*}
-	We solve the multipliers and have:
-	\begin{align*}
-		\rho_{\mathcal{c}} = &\frac{1}{ 2 \pi \sigma_q \sigma_p} e^{-\frac{(q-\mu_q)^2}{2\sigma_q^2}} e^{-\frac{(p-\mu_p)^2}{2\sigma_p^2}} \\
-		I_0 = &\ln (2\pi e\sigma_q\sigma_p) \\
-		\sigma_q \sigma_p &= \frac{e^{I_0}}{2 \pi e}
-	\end{align*}
-	This shows that the gaussian minimizes the spread at fixed entropy, therefore all other distributions must have a larger or equal spread. 
-\end{proof}
-
-\begin{remark}
-	Note that the inverse argument does not work: if we have a bound on the uncertainty, we cannot say anything about the entropy.
-All distributions with higher entropy will satisfy a higher bound, therefore the entropy is arbitrarily high.
-Moreover, we can find distributions with low entropy but with arbitrarily high spread.
-Therefore a bound on the uncertainty still allows any value for entropy.
-\end{remark}
-
 \bibliography{bibliography}

### 2021-11-15 Articles/2021-ReversePhysics/Submissions/Nature/ReversePhysics.tex
ReversePhysics: Changing the abstract

@@
 
-\title{Reverse Physics}
-\author{Gabriele Carcassi, Christine A. Aidala}}
+\title{Reverse Physics: From Laws to Physical Assumptions}
+\author{Gabriele Carcassi, Christine A. Aidala}
 \affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
@@
 \begin{abstract}
-We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics.
-The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program.
-This allows us to identify physical starting points (as opposed to mathematical ones), compare their deductive power, and find those that can better serve as a foundation for our theories.
+To answer foundational questions in physics, physicists turn more and more to abstract advanced mathematics, even though its physical significance may not be immediately clear.
+What if we started to borrow ideas and approaches, with appropriate modifications, from the foundations of mathematics?
+In this paper we explore this route.
+In reverse mathematics\cite{friedman1976systems,simpson2017reverse,stillwellreverse} one starts from theorems and finds the minimum set of axioms required for their derivation.
+In reverse physics we want to start from laws or more specific results, and find the physical concepts and starting points that recover them.
+We want to understand what physical results are implied by which physical assumptions.
+As an example of the technique, we will see six different characterizations of classical mechanics, show that the uncertainty principle depends only on the entropy bound on pure states and recast the third law of thermodynamics in terms of the entropy of an empty system.
+We believe the approach can provide greater insights into both current and new physical theories, put the physical concepts at the forefront of the discussion and provide a more unified view of physics by highlighting common patterns and ideas across different physical theories.
 \end{abstract}

### 2021-11-18 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
ClassicalFail: changes after discussions

@@
 We present the idea that, contrary to the established view, classical mechanics fails for conceptual reasons, not merely because it does not match experimental evidence.
-The notion of a classical system stems from two underlying assumptions, perfect isolation and infinitesimal divisibility, that are in contradiction: at some point we will reach a scale for which the parts can no longer be assumed to be isolated.
+The notion of a classical system stems from two underlying assumptions, perfect isolation and infinitesimal divisibility, that can never be physically realized at the same time: at some point we will reach a scale for which the parts can no longer be assumed to be isolated.
 This result defines more cleanly the realm of applicability of classical mechanics, shows what problems quantum mechanics addresses and gives hints as to what other problems will need to be solved by a more general theory.
@@
 	\item The interaction across the boundary participates in the definition of the system
-	\item Perfect characterization of a system requires perfect isolation
+	\item Perfect characterization of the values of all properties of a system requires perfect isolation
 	\item As we get to smaller and smaller parts, perfect isolation becomes untenable
@@
 A cannonball at 5000 Kelvin does not exist, or at least not for long.
-Conceptually, we can think of a property as being intrinsic to the system only if the interaction with the environment leaves it, at least to a good approximation, undisturbed.
+The state of a system, therefore, does not depend only on the object we are studying, but also on its context.
+
+The ability to model a property of a system as a real valued quantity depends on the assumption that the environment interaction can be, at least in some cases, made to be negligible.
+We can, for example, remove the effect of the air on the speck of dust by putting it in a vacuum.
+It is that assumption that allows us to model the state of a classical system as a set of infinitely precise values.
+Unfortunately, this model is so ingrained that, even when the assumption fails, we are tempted to still think that there is a \emph{true} value, even though it is not experimentally accessible.
+But if the assumption fails, the model is no longer justifiable and a new more complex model, based on new assumptions, need to take its place.\footnote{As an analogy, to the naked eye a star looks like a point in the sky and can be assumed to be as such.
+Even with several order of magnification, the assumption will hold.
+However, at some point one reaches a magnification such that the star becomes bigger and the assumption break down.
+At that point, you can't describe the star as just a single point: it's a more complicated object.
+In the same way, if we can't assume the environment interaction can be removed, the state is not a set of infinitely precise values over which we add a statistical interaction: it has to be something more complicated.}
+
+In general, we can think of a property as being intrinsic to the system itself only if the interaction with the environment leaves it, at least to a good approximation, undisturbed.
 If it is not left undisturbed, then it is no longer a property of the system itself: it is a shared property between the system and the environment because the nature of both comes into play in its definition.\footnote{CITE: maybe someone in complex systems, emergent properties, ...
-make similar claims}
-
-While we are used to think this way for thermodynamic quantities, like temperature or pressure, or statistical quantities, like average energy or mean free path, we typically think of mechanical properties, like position and velocity, differently.
+make similar claims} While we are used to think this way for thermodynamic quantities, like temperature or pressure, or statistical quantities, like average energy or mean free path, we typically think of mechanical properties, like position and velocity, differently.
 These, we think, are always intrinsic to the system.
@@
 What happens is that, in the vast number of cases, we can sufficiently isolate the system so that the effect of the environment on the motion of the body is negligible.
-We can take a speck of dust, put it in a vacuum, and the position is now much more stably definable.
-Therefore, when studying the speck in a fluid, the motion is modeled as fluctuations of an infinitely precise quantity because, at least conceptually, we could instantly remove the air and determine the position of the speck in isolation.
 But, realistically, we can only go so far.
@@
 For large scale objects, this is not a problem: small perturbations will have small effects which can, at least at a certain scale, be neglected.
-But this is an idealization.
-The infinite precision of a real number is valid only insofar that the idealization holds.
+Assuming perfect isolation by itself, therefore, is a reasonable approximation in range of cases.
 
-Assuming perfect isolation by itself is not an issue.
 However, classical mechanics also assumes, implicitly, that this perfect isolation holds for objects of all sizes.
@@
 The theory implicitly assumes all these objects can be prepared, manipulated and detected independently enough that their state can be perfectly defined.
-This is the contradiction inherent to classical mechanics: perfect isolation at all scales.
+This is the inherent problem with classical mechanics: perfect isolation at all scales.
+The issue is that perfect isolation itself is a simplifying assumption, and as we go to smaller and smaller scales the assumption becomes less and less tenable.
 
@@
 
-No matter our definition, then, the entropy of a single microstate is minus infinity. This certainly does not sound terribly reassuring... but is it a problem?
+No matter our definition, then, the entropy of a single microstate is minus infinity.
+This certainly does not sound terribly reassuring...
+but is it a problem?
+In information theory, for example, entropy on the continuum is seen by many as a relative concept.
+Yet, information entropy is not exactly the same as thermodynamic entropy.\footnote{The Shannon entropy corresponds to the Gibbs entropy, and therefore the thermodynamic entropy, only when computed over canonical coordinates (such as position and conjugate momentum).} So, does it make sense for the thermodynamic entropy to be minus infinity?
 
@@
 The idea is that a perfect crystal at zero Kelvin only has one possible configuration: it has one possible microstate.
-But we have seen that the entropy of a single microstate is minus infinity, not zero.
+But we have seen that the entropy of a single classical microstate is minus infinity, not zero.
 Therefore the result is in contradiction.
@@
 This removes all the inconsistencies with thermodynamics, statistical mechanics and information theory.
-Conceptually, it signifies there is always a background interaction with the environment that cannot be fully eliminated.
+This feature of quantum mechanics can be understood as assuming that there is always a background interaction with the environment that cannot be fully eliminated.
 
 If there is interaction at the system/environment boundary, then the properties of the system can only be defined up to the statistical fluctuations imposed by the interaction.
-In fact, in quantum mechanics the probabilistic description is baked into the very notion of state, it is not something that can be removed like in classical statistical mechanics.
+This is in line with quantum states giving an inherent probabilistic description; it is not something that can be removed like in classical statistical mechanics.
 

### 2021-11-23 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: draft of the main argument

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\RequirePackage{amsthm}
\RequirePackage{amsmath}
\RequirePackage{amssymb}
\RequirePackage{enumitem}
\setlist{noitemsep}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\newcommand\partitle[1]{\textsc{#1}.}

\begin{document}

\title{On the reality of the quantum state once again}
\author{Gabriele Carcassi, Andrea Oldofredi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics. The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program. This allows us to identify physical starting points (as opposed to mathematical ones), compare their deductive power, and find those that can better serve as a foundation for our theories.
\end{abstract}

\maketitle

\section{Introduction}
%(mezza pagina)
%- Status of field and impact of HS
%- Claim that HS does not work for quantum epistemic cases, therefore invalidating all results

\section{Summary of the model}
%(una pagina)
%- Ontic/supplemented/epistemic

\section{Quantum epistemic states}
Now that we have presented the model, we want to study the space of epistemic states. To make the discussion more concrete, we will assume we have a $\psi$-complete ontological model of a single qubit representing the direction of spin.

Suppose we have two preparation devices, one that prepares z+ and another that prepares z-, and we toss a fair coin to choose which one to use. This results into an epistemic state with equal distribution on z+ and z-. Now suppose we have two additional preparation devices, this time for x+ and x-, and we again toss a fair coin. This time we have an epistemic state with equal distribution on x+ and x-. Note that the epistemic distributions do not overlap. Therefore, according to the HS categorization, these correspond to two ontologically distinct states, making the preparation axis an ontological property of the system, regardless of the orientation. However, in terms of density matrixes both these situations would not just overlap: they would be represented by the same maximally mixed state. Hence there seems to be a mismatch between the type of epistemic knowledge one can express in quantum mechanics through mixed states and the type of epistemic knowledge described in the HS model. This needs to be investigated further.

Let us first characterize the extent of the discepancy. In HS, the space of epistemic states $E_{HS}$ is the set of probability distributions over pure states, which is an infinite dimensional space. This is a direct consequence of defining the epistemic states as a distribution $P(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability to each ontological state $\lambda$.\footnote{TODO: HS seems to be very vague about probability and probability density.} In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold. Looking at the difference in dimension, the discrepancy is not minor.

Note that the space of pure states $\Lambda$ is the same in both cases, one state for each direction of spin. Also note that all epistemic states in both spaces can be reached by mixture of pure states. The difference is that, in the quantum case, we have an ``aliasing'': different mixtures will give the same mixed states. Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{Q}$ that allows us to go from HS epistemic states to density matrixes. This map is surjective but not injective. Therefore $E_{Q}$ can be seen as a set of equivalence classes of $E_{HS}$. The key question, then, is the following: is the map $\iota$ sufficiently well behaved that we can understand the epistemic content of $E_{Q}$ based on the epistemic content of $E_{HS}$? Or is it so hopelessy broken that there is no connection between the two?

Ideally, one would like to say is that $E_{HS}$ represents all possible epistemic states that include the knowledge about preparation, while $E_{Q}$ represents only those distinguishable through measurement. For example, one may want to say that the axis of preparation is, in principle, an ontologically distinct property, but this is not experimentally distinguishable. Therefore distinct epistemic states for each direction are projected by $\iota$ onto the same epistemic state of maximum uncertainty. This sounds very plausible and appealing, but it turns out this is not at all the case. To show this we will concentrate on a single aspect about the map: how information entropy transforms under the map. While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is sufficient to claim that the epistemic structures of $E_{HS}$ and $E_{Q}$ are different.

Since every pure state in quantum mechanics has zero entropy, we will assume that this is the case for all pure states in $E_{HS}$. Let us consider three uniform distributions over a sphere with support given by:
\begin{description}
	\item[A] the poles, z+ and z-
	\item[B] the equator, all states on the x/y planes
	\item[C] the whole sphere, all points
\end{description}
Since we want the entropy for a single microstate to be zero, the entropy for A will be 1 (uniform distribution between two discrete cases). Since $\iota(A)$ will be the maximally mixed state, which also corresponds to entropy 1. So far so good: $\iota$ maps entropy 1 to entropy 1! However, note that A is a measure zero set of B, which is a measure zero set of C. The entropy of C is therefore infinite with respect to B which is infinite with respect to A. On the other hand, A, B and C will all correspond to the same maximally mixed state, therefore $\iota$ collapses together epistemic states at different level of infinity. Note that, in this case, $\iota$ picks the minimum entropy between the equivalent epistemic states. This is inconsistent with the idea that $E_{Q}$ represents states where knowledge is lost: that would pick the maximum possible entropy.

Now consider another uniform distribution over the support:
\begin{description}
	\item[D] the north pole and a point on the equator, z+ and x+.
\end{description}
This is yet another uniform distribution between two discrete cases, therefore in $E_{HS}$ the entropy is 1. However, $\iota(D)$ is not the maximally mixed state and therefore the entropy is less than 1. This means that epistemic states that are isoentropic in $E_{HS}$ are not isoentropic $E_{Q}$. If we think as $E_{HS}$ and $E_{Q}$ as partially ordered by information entropy, then $\iota$ is not an order preserving map. Therefore the epistemic structures represented by $E_{HS}$ and $E_{Q}$ are not isomorphic.

The seemingly innocuous assumption that $P(\lambda|P)$ is well defined for all preparations $P$ and for all states $\lambda$ imposes a classical probability space which follows the rules of classical information theory, and prevents the state space $\Lambda$ to support an epistemic structure compatible with quantum theory. This suggests an obvious follow up: what if we start by assuming that the set of pure states $\Lambda$ supports the correct epistemic structure, the one associated with density matrices, and use the same definitions of $\psi$-ontic and $\psi$-epistemic?

The difficulty here is that support and overlaps are not a well defined concept for density matrices. Yet, there are two different properties that non-overlapping functions have in classical probability theory. Two distribution $\rho_1$ and $\rho_2$ are non-overlapping if and only if:
\begin{enumerate}
	\item their product $\rho_1 \rho_2 = 0$ is zero
	\item the probability of being in the support of the second given the first is zero.
\end{enumerate}
In quantum mechanics, if we consider two pure states $\psi_1$ and $\psi_2$, they lead to the following conditions:
\begin{enumerate}
	\item $|\psi_1><\psi_1|\psi_2><\psi_2| = 0$
	\item $<\psi_2|\psi_1><\psi_1|\psi_2> = 0$.
\end{enumerate}
Both of these are satisfied if and only if $<\psi_1|\psi_2>=0$: the states are non-overlapping if and only if they are orthogonal. This is, not suprisingly, exactly what happens in classical probability: two distributions are non-overlapping if and only if they are orthogonal with respect to the inner product $\int \rho_1 \rho_2 d\lambda$.\footnote{We could also define the overlap in terms of the entropy of the mixture and we would find the same result.}

Therefore we have the following result: if we want a set of pure states that are able to support the epistemic structure of quantum mechanics, they must be ``overlapping'' and therefore they cannot be take to be ontic. The reader is no less entitled to conclude that quantum mechanics does not allow $\psi$-ontic interpretations from this result than he would be entitled to conclude that quantum mechanics does not allow $\psi$-epistemic interpretations from PBR. He would also be entitled to conclude that the categorization is flawed.

%- Try to apply their model to a purely mixed state of spin 1/2
%- Show that you can't have a single epistemic distribution
%-> Failure of model

\section{Discussion}

Let us now step back a moment. While the math is clear enough, we still need to understand what is going on at a conceptual level. Why does assuming $p(\lambda|P)$ lead to an epistemic structure incompatible with the quantum one? To answer this question, let us review how probability spaces work in both classical and quantum mechanics.

A ``classical'' probability space is made of three objects: a sample space $\Omega$ which represents all possible cases (e.g. all points in phase-space for a classical system); a $\sigma$-algebra $\Sigma_\Omega$ over $\Omega$ which represents all statements of interest (e.g. ``the position is between 2 and 3 meters''); a measure $\mu : \Sigma_\Omega \to \mathbb{R}$ that associates each statement with a probability. Note how we have a single probability distribution over the whole space.

In quantum mechanics we start with a similar structure: a set of states $\mathcal{H}$ which represents all possible cases (i.e. the Hilbert space) and a $\sigma$-algebra $\Sigma_{\mathcal{H}}$ that represents all statements of interests. This is where the similarity ends. To specify a mixed state, a ``quantum distribution'', we use a density matrix $\rho : \mathcal{H} \to \mathcal{H}$. To get a probability distribution, we need to specify an observable $O$. The observable $O$ identifies an orthogonal basis and we take the sub-algebra $\Sigma_O \subset \Sigma_{\mathcal{H}}$ defined on those states. We can then associate $\Sigma_O$ with a measure $\mu_O : \Sigma_O \to \mathbb{R}$, which is therefore not on the whole space $\Sigma_{\mathcal{H}}$.

This procedure is repeated for all observables. The probability measure of each sums to one over the respective sub-algebra, over the respective basis. There are infinitely many observables (e.g. each direction of spin) whose bases are disjoint sets. If we added all measures over the entire $\sigma$-algebra (e.g. all probabilities over all directions of spin) the sum would be infinite. In other words: we do not have a single probability distribution on a single $\sigma$-algebra. For each observable, we a probability distribution defined on a particular sub-algebra\footnote{This is exactly why some rules of classical probability are not allowed in quantum probability: we can only combine statements within the same sub-algebra and not across different sub-algebras.} and this is what gives rise to contextuality. Before getting a probability distribution we need to specify a context, a sub-algebra, a basis.

We may be tempted to think that we have one context before and one after the measurement, but this does not quite work. Each context is needed, for example, to define the expectation value of the related observables. It is by measuring those quantities over an ensemble of identically prepared system (i.e. quantum tomography) that we can reconstruct the initial state. Therefore we need to be able to say that all distributions over all contexts are fixed by the preparation. The measurement process can only measure one without perturbing the other.

Yet, at each time there is a special context. If we take the $z+$, the probability distribution along that direction will the sharpest while the probability distribution along a perpendicular direction will be uniform. Therefore there is a context in which the entropy is minimal, there is a context that gives us the most information about the system. This is the basis over which we calculate the entropy of a mixed state, this is the basis in which the density matrix diagonalizes.

Note that the minimal entropy context changes over mixes. For an equal mix of $z+$ and $x+$, the minimal entropy distribution is neither over $Z$ or $X$. This is why the epistemic structure of contextual and non-contextual models is so different: classical mixtures affect the context.

To sum up, assuming $p(\lambda|P)$ exists means assuming there is a single probability space, a single $\sigma$-algebra, a single set of propositions, to which all probabilities are assigned. This is exactly assuming non-contextuality. To allow for contextuality we should assume that for each observable $O$ we have a probability $p_O(\lambda_O|P)$ of having prepared one of the eigenstates, such that $\sum_{\lambda_O} p_O(\lambda_O|P) = 1$. However, the sum over all states $\sum_{\lambda} p_{O(\lambda)}(\lambda|P) = + \infty$ diverges and is not a probability measure. If it didn't diverge, then we could renormalize and put a probability measure over the whole space, making the model non-contextual. Mathematically, the probability space is not Kolmogorov: probability of countable sequences does not always converge.\footnote{In our larger project, we are trying to understand what are the correct axioms for such space with mixed success.}

% (mezza pagina)
%- Non-contestuality of HS
%- PBR and other results

\section{Conclusion}

\bibliography{bibliography}

\end{document}

### 2021-11-24 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: first pass at the main argument

@@
 
-Let us now step back a moment.
-While the math is clear enough, we still need to understand what is going on at a conceptual level.
-Why does assuming $p(\lambda|P)$ lead to an epistemic structure incompatible with the quantum one?
-To answer this question, let us review how probability spaces work in both classical and quantum mechanics.
-
-A ``classical'' probability space is made of three objects: a sample space $\Omega$ which represents all possible cases (e.g.
-all points in phase-space for a classical system); a $\sigma$-algebra $\Sigma_\Omega$ over $\Omega$ which represents all statements of interest (e.g.
-``the position is between 2 and 3 meters''); a measure $\mu : \Sigma_\Omega \to \mathbb{R}$ that associates each statement with a probability.
-Note how we have a single probability distribution over the whole space.
-
-In quantum mechanics we start with a similar structure: a set of states $\mathcal{H}$ which represents all possible cases (i.e.
-the Hilbert space) and a $\sigma$-algebra $\Sigma_{\mathcal{H}}$ that represents all statements of interests.
-This is where the similarity ends.
-To specify a mixed state, a ``quantum distribution'', we use a density matrix $\rho : \mathcal{H} \to \mathcal{H}$.
-To get a probability distribution, we need to specify an observable $O$.
-The observable $O$ identifies an orthogonal basis and we take the sub-algebra $\Sigma_O \subset \Sigma_{\mathcal{H}}$ defined on those states.
-We can then associate $\Sigma_O$ with a measure $\mu_O : \Sigma_O \to \mathbb{R}$, which is therefore not on the whole space $\Sigma_{\mathcal{H}}$.
-
-This procedure is repeated for all observables.
-The probability measure of each sums to one over the respective sub-algebra, over the respective basis.
-There are infinitely many observables (e.g.
-each direction of spin) whose bases are disjoint sets.
-If we added all measures over the entire $\sigma$-algebra (e.g.
-all probabilities over all directions of spin) the sum would be infinite.
-In other words: we do not have a single probability distribution on a single $\sigma$-algebra.
-For each observable, we a probability distribution defined on a particular sub-algebra\footnote{This is exactly why some rules of classical probability are not allowed in quantum probability: we can only combine statements within the same sub-algebra and not across different sub-algebras.} and this is what gives rise to contextuality.
-Before getting a probability distribution we need to specify a context, a sub-algebra, a basis.
-
-We may be tempted to think that we have one context before and one after the measurement, but this does not quite work.
-Each context is needed, for example, to define the expectation value of the related observables.
-It is by measuring those quantities over an ensemble of identically prepared system (i.e.
-quantum tomography) that we can reconstruct the initial state.
-Therefore we need to be able to say that all distributions over all contexts are fixed by the preparation.
-The measurement process can only measure one without perturbing the other.
-
-Yet, at each time there is a special context.
-If we take the $z+$, the probability distribution along that direction will the sharpest while the probability distribution along a perpendicular direction will be uniform.
-Therefore there is a context in which the entropy is minimal, there is a context that gives us the most information about the system.
-This is the basis over which we calculate the entropy of a mixed state, this is the basis in which the density matrix diagonalizes.
-
-Note that the minimal entropy context changes over mixes.
-For an equal mix of $z+$ and $x+$, the minimal entropy distribution is neither over $Z$ or $X$.
-This is why the epistemic structure of contextual and non-contextual models is so different: classical mixtures affect the context.
-
-To sum up, assuming $p(\lambda|P)$ exists means assuming there is a single probability space, a single $\sigma$-algebra, a single set of propositions, to which all probabilities are assigned.
-This is exactly assuming non-contextuality.
-To allow for contextuality we should assume that for each observable $O$ we have a probability $p_O(\lambda_O|P)$ of having prepared one of the eigenstates, such that $\sum_{\lambda_O} p_O(\lambda_O|P) = 1$.
-However, the sum over all states $\sum_{\lambda} p_{O(\lambda)}(\lambda|P) = + \infty$ diverges and is not a probability measure.
-If it didn't diverge, then we could renormalize and put a probability measure over the whole space, making the model non-contextual.
-Mathematically, the probability space is not Kolmogorov: probability of countable sequences does not always converge.\footnote{In our larger project, we are trying to understand what are the correct axioms for such space with mixed success.}
-
 % (mezza pagina)
@@
 
+\section*{Supplementary material}
+
+A ``classical'' probability space is made of three objects: a sample space $\Omega$ which represents all possible cases; a $\sigma$-algebra $\Sigma_\Omega$ over $\Omega$ called event space, where each event is represented by a set of outcomes in the sample space; a measure $\mu : \Sigma_\Omega \to \mathbb{R}$ that associates each statement with a probability.
+Note how we have a single probability distribution over the whole space.
+
+To make this concrete, let us represent the space of all possible directions using the surface of a sphere.
+The set of all possible points on the surface is our sample space $\Omega$.
+This space is endowed with a topology and therefore a Borel algebra $\Sigma_\Omega$, which is the
+
+In quantum mechanics we start with a similar structure: a set of states $\mathcal{H}$ which represents all possible cases (i.e.
+the Hilbert space) and a $\sigma$-algebra $\Sigma_{\mathcal{H}}$ that represents all statements of interests.
+This is where the similarity ends.
+To specify a ``quantum distribution'' we use a density matrix $\rho : \mathcal{H} \to \mathcal{H}$.
+To get a probability distribution, we need to first specify an observable $O$.
+The observable $O$ identifies a basis $B_O \subset \mathcal{H}$ and we take the sub-algebra $\Sigma_O \subset \Sigma_{\mathcal{H}}$ defined on those states.
+We can then associate $\Sigma_O$ with a measure $\mu_O : \Sigma_O \to \mathbb{R}$, which is therefore not on the whole space $\Sigma_{\mathcal{H}}$.
+This different setup is what is responsible for the differences between classical and quantum probability.
+We highlight here some of the main ones.
+
+\textbf{Contextuality}.
+Before getting a probability distribution we need to specify an observable, a basis, a sub-algebra: we need to specify a context in which the probability distribution is specified.
+This is what we refer to as contextuality.
+For example
+
+\textbf{Forbidden across-context operations}.
+While we can perform all countable logical operations within the $\sigma$-algebra $\Sigma_{O}$ of each context, if we try to combine events across context, through a logical AND or logical OR, we will inevitably go outside both contexts.
+The event still exists in the overall $\sigma$-algebra $\Sigma_{\mathcal{H}}$, so we can talk about it from a strictly logical perspective.
+Yet, it may not exist in any context and cannot be directly given a probability.
+
+\textbf{Failure of countable additivity}.
+Another problem is that, in quantum mechanics, we have probability distributions that fail to be countably additive.
+An eigenstate of momentum, for example, is invariant under translation which necessary means the probability distribution over position must be uniform.
+Yet, a uniform distribution over an infinite range is not something expressible through Kolmogorov probability as it violates countable additivity.
+Each finite interval, in fact, must necessarily correspond to zero probability yet the full range must correspond to probability one.
+
 \bibliography{bibliography}

### 2021-12-01 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: another pass and added discussion

--- 2021-11-23 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
@@
 
-Let us now step back a moment.
-While the math is clear enough, we still need to understand what is going on at a conceptual level.
-Why does assuming $p(\lambda|P)$ lead to an epistemic structure incompatible with the quantum one?
-To answer this question, let us review how probability spaces work in both classical and quantum mechanics.
+Let us now step back a moment. Why does assuming $p(\lambda|P)$ lead to an epistemic structure incompatible with the quantum one?
+
+It is widely understood that one feature of quantum mechanics is non-contextuality: a measurement cannot be understood as revealing pre-existing values.
+We may be tempted to think that we have one context before and one after the measurement.
+What we are seeing that this does not work because we a dual effect on preparations.
+As we cannot understand quantum measurements as selecting from a single joint probability distribution, we cannot understand quantum preparations as constructing a single joint probability distribution.
+To fully understand how this works, let us review how how probability spaces work in both classical and quantum mechanics.
 
@@
 
-This procedure is repeated for all observables.
-The probability measure of each sums to one over the respective sub-algebra, over the respective basis.
-There are infinitely many observables (e.g.
-each direction of spin) whose bases are disjoint sets.
-If we added all measures over the entire $\sigma$-algebra (e.g.
-all probabilities over all directions of spin) the sum would be infinite.
-In other words: we do not have a single probability distribution on a single $\sigma$-algebra.
-For each observable, we a probability distribution defined on a particular sub-algebra\footnote{This is exactly why some rules of classical probability are not allowed in quantum probability: we can only combine statements within the same sub-algebra and not across different sub-algebras.} and this is what gives rise to contextuality.
-Before getting a probability distribution we need to specify a context, a sub-algebra, a basis.
+Mathematically, what we call context is a $\sigma$-algebra: a set of propositions upon which we assign probabilities.
+This tells us why we cannot simply use classical rules in quantum mechanics.
+Classical mechanics has only one context, so we can always mix and match statements however we please.
+In quantum mechanics, propositions about different observables live in different $\sigma$-algebras, in different probability spaces, and therefore cannot be combined.
 
-We may be tempted to think that we have one context before and one after the measurement, but this does not quite work.
-Each context is needed, for example, to define the expectation value of the related observables.
-It is by measuring those quantities over an ensemble of identically prepared system (i.e.
-quantum tomography) that we can reconstruct the initial state.
-Therefore we need to be able to say that all distributions over all contexts are fixed by the preparation.
-The measurement process can only measure one without perturbing the other.
+One may think that we can get away by assuming that, at each time, there is a ``true'' context.
+And this seems to work in some cases.
+Given any density matrix $\rho$, there exist a (non-unique) privileged context: this is the one in which we can express $\rho = \sum p_i |\psi_i > < \psi_i|$ as the mix of orthogonal pure states.
+If $\rho$ is itself a pure state, any basis that includes it will be a privileged context.
+This privileged context provides us the probability distribution over which the information entropy is calculated.
+After a measurement, the privileged context changes to one that includes the eigenstates of the chosen observable.
+However, this does not work in general.
 
-Yet, at each time there is a special context.
-If we take the $z+$, the probability distribution along that direction will the sharpest while the probability distribution along a perpendicular direction will be uniform.
-Therefore there is a context in which the entropy is minimal, there is a context that gives us the most information about the system.
-This is the basis over which we calculate the entropy of a mixed state, this is the basis in which the density matrix diagonalizes.
+Suppose we start with two density matrixes $\rho_1$ and $\rho_2$ with different eigenstates.
+We then create the mixture $\rho = p_1 \rho_1 + p_2 \rho_2$.
+The contexts of the mixture will not be the same as the original contexts.
+It will, in fact, even depend on $p_1$, $p_2$.
+That is, when mixing preparations we are not simply combining two probability distribution over the same context, like in classical mechanics, but we are combining the contexts themselves.
+Non-contextuality does not affect only measurement, but, dually, preparations as well.
+And this affect how information entropy is calculated and the overall epistemic structure.
 
-Note that the minimal entropy context changes over mixes.
-For an equal mix of $z+$ and $x+$, the minimal entropy distribution is neither over $Z$ or $X$.
-This is why the epistemic structure of contextual and non-contextual models is so different: classical mixtures affect the context.
+% TODO: mention tomography?
 
@@
 This is exactly assuming non-contextuality.
+While this can be made to work for pure state to pure state transitions during measurement, mixing preparations leads to an epistemic structure that is fundamentally non-contextual, leading to inconsistencies with quantum mechanics.
 To allow for contextuality we should assume that for each observable $O$ we have a probability $p_O(\lambda_O|P)$ of having prepared one of the eigenstates, such that $\sum_{\lambda_O} p_O(\lambda_O|P) = 1$.
-However, the sum over all states $\sum_{\lambda} p_{O(\lambda)}(\lambda|P) = + \infty$ diverges and is not a probability measure.
-If it didn't diverge, then we could renormalize and put a probability measure over the whole space, making the model non-contextual.
-Mathematically, the probability space is not Kolmogorov: probability of countable sequences does not always converge.\footnote{In our larger project, we are trying to understand what are the correct axioms for such space with mixed success.}
+The entropy of the preparation would be the lowest entropy across all contexts.
 
@@
 
+\section*{Supplementary material}
+
+TODO: do we need to bump anything here?
+
 \bibliography{bibliography}

### 2021-12-07 Overviews/Informational granularity/InformationGranularity.tex
InformationGranularity: added a couple examples

@@
 
+\section{Spaces we need to be able to capture}
+
+\textbf{Discrete spaces.} Suppose we have a domain that counts how many objects are there (e.g.
+pieces of fruit in a bowl, particles in a box or a piece of material).
+The problem is, in principle, decidable; the theoretical domain is the power set of the natural number.
+The measure will be the counting measure.
+Therefore, we will have that all possibilities are equigranular.
+This is enough to recover the counting measure.
+Infinite sets may or may not be considered comparable.
+TOCHECK: We can add the comparability of mod classes.
+
+\textbf{Riemannian spaces}.
+Let $X$ me a Riemannian manifold of dimension $n$.
+Let $S^k$ be the set of closed submanifold of dimension $k$ (i.e.
+the submanifold that are closed sets).
+We set $\sigma_1 \sfiner \sigma_2$ for all $\sigma_1 \in S^i$, $\sigma_2 \in S^k$ such that $i < j$.
+If $i = j$, then we set $\sigma_1 \finer \sigma_2$ if the area/volume associated to $\sigma_1$ is less or equal to the one associated to $\sigma_2$.
+For manifold of dimension zero, we use the counting measure.
+TOCHECK: The fineness releationship so defined are enough to recover the geometric structure.
+
+\textbf{Synmplectic spaces}.
+Let $X$ me a symplectic manifold of dimension $sn$.
+Let $S^{2k}$ be the set of closed submanifold of dimension $2k$ (i.e.
+the submanifold that are closed sets).
+We set $\sigma_1 \sfiner \sigma_2$ for all $\sigma_1 \in S^{2i}$, $\sigma_2 \in S^{2k}$ such that $i < j$.
+If $i = j$, then we set $\sigma_1 \finer \sigma_2$ if the area/volume associated to $\sigma_1$ is less or equal to the one associated to $\sigma_2$.
+For manifold of dimension zero, we use the counting measure.
+TOCHECK: The fineness releationship so defined are enough to recover the geometric structure.
+
 \section{Insightful failures}
@@
 
+\subsection{Open set comparability}
+
+\textbf{Takeaway: we can't assume all verifiable statements (open sets) are comparable.}
+
+There are two main reasons that assuming all verifiable statements are comparable does not work.
+The first is physical dimension.
+Consider the two statements \statement{the position of the ball is between 1 and 2 m} and \statement{the velocity of the ball is between 1 and 2 m/s}.
+These cannot be compared.
+
+The second reason is because of how infinity works.
+Consider that domain formed by all statements regarding counting the number of pieces of fruit in a bowl.
+This would be equivalent to the power set of the natural numbers, since the domain is decidable (i.e.
+all statements are clopen).
+Now suppose that we say all verifiable statements are comparable.
+This means the fineness is a total/linear preorder.
+TO CHECK: A known result in mathematical logic is a total order on the power set of infinitely many elements does not exist.
+
 \bibliography{bibliography}

### 2021-12-09 Overviews/Informational granularity/InformationGranularity.tex
ImformationGranularity: added another example

@@
 
-\section{Spaces we need to be able to capture}
+\subsection{Example of lack of equigranular or comparable possibilities}
+
+Consider the case of a balanced die.
+It will have six equigranular possibilities.
+Consider the case of an unbalanced die.
+It will have six possibilities, but they will not be equigranular.
+Consider the case where you may have a balanced or unbalanced die.
+It is not clear whether these cases should be equigranular or not.
+Therefore we should leave that as not comparable.
+In this case, the possibilities are twelve divided into two groups.
+Two possibilities with the group will be comparable but not across groups.
+
+\subsection{Spaces we need to be able to capture}
 

### 2021-12-09 BareMinima/Template/Template.tex
BM-Template: template draft

--- 2021-05-19 BareMinima/SetTheory/BareMinimumSetTheory.tex
@@
 
-\title{Bare minimum: set theory}
+\title{Bare minimum: SUBJECT}
 
@@
 \begin{abstract}
-This note presents a condensed summary of set theory which can function as a crash course, refresher and/or reference.
+This note presents a condensed summary of SUBJECT which can function as a crash course, refresher and/or reference.
 Bare minima are meant to give a rough overview, by no means complete, of the subject to the intellectually curious, particularly in the context of foundational questions in physics.
-This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
 \end{abstract}
@@
 
-Set theory is important for those working on the Assumptions of Physics for at least three reasons.
-First, it is a foundational framework in mathematics.
-Second, it showcases a successful attempt at concept generalization.
-Third, it shows how a formal system is bootstrapped.
+Brief overview
 
-There are two rough branches of set theory: naive and axiomatic.
-Naive set theory is built on intuitive concepts, and as such is not fully formalized and is open to potential paradoxes.
-Axiomatic set theory provides a fully formalized axiomatic system that aims to close those problems.
-We will start with naive set theory, which is the best setting for defining all the basic notions and getting a sense of how the framework works.
-Then we will turn our attention to axiomatic set theory to get a sense of what problems it solves and how, and whether those solutions mesh with foundational goals in physics.
+\emph{This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.}
 
-For more details on set theory, see for example \cite{settheory-tutorial, pinter2014book}.
+\section{A topic}
 
-\section{Naive set theory}
+Introductory comment of an informal nature.
 
-We take sets and elements as primitive objects, with no formal definition.
-A \textbf{set} $A$ is a collection of elements.
-If an element $e$ belongs to the set we say $e$ \textbf{is in} $A$, noted $ e
-\in A$.
-A set $A$ can be defined by listing the elements, noted $A = \{ e_1, e_2, e_3 \}$.
-A set $A$ can be defined by specifying how to build the elements through a rule (i.e.
-predicate) $P(e)$, and is noted $A = \{e \, | \,  P(e) \}$.
-The rule may include symbols like for all $\forall$, exists $\exists$, logical connectors AND $\AND$, OR $\OR$ and $\NOT$.
+\subsection{A subtopic}
 
-\begin{defn}[Common sets]
-	 The \textbf{empty set}
-	 \marginleft{Common sets: $\emptyset$, $\{a\}$, $\mathbb{N}$, $\mathbb{Z}$, $\mathbb{Q}$, $\mathbb{R}$, $\mathbb{C}$}
-	  $\emptyset$ is the set with no elements. A \textbf{singleton} is the set of a single value $\{a\}$.
-	 The \textbf{set of natural numbers} $\mathbb{N} = \{ 0, 1, 2, ...
-\}$.
-The \textbf{set of positive natural numbers} $\mathbb{N}^+ = \{n \in \mathbb{N} \, | \, n > 0\}$.
-The \textbf{set of integer numbers} $\mathbb{Z} = \{ ...
-, -2, -1, 0, +1, +2, ...\}$.
-The \textbf{set of rational numbers} $\mathbb{Q}$.
-The \textbf{set of real numbers} $\mathbb{R}$.
-The \textbf{set of complex numbers} $\mathbb{C}$.
-A \textbf{universe} $U$ is the set that, within a specific context, contains all entities under study.
+\begin{defn}[Pop of a set - simple definition]
+	The pop of a set $A$ \marginleft{Pop} is the curb of its power set.
 \end{defn}
 
-\subsection{Basic definitions}
+\begin{defn}[Foo sets - definition with properties]
+	We say a set $A$ is \textbf{foo}
+	\marginleft{Foo: $\circledcirc$}, noted $\circledcirc A$, if it satisfies the following properties:
+	\begin{itemize}
+		\item \textbf{Shirted}: the pop of $A$ is shirted
+		\item \textbf{Intarred}: the pop of $A$ is a subset of the tar of $A$
+	\end{itemize}
+\end{defn}
+
+\begin{defn}[Foo operations - multiple definitions]
+	We define the following operations on foo sets:
+	\begin{description}
+		\item[Credit.] Noted $A \veebar B$, the credit of $A$ and $B$ is the smallest foo set that contains.
+		\item[Change.] Noted $A \barwedge B$, the change of $A$ and $B$ is the biggest foo set that is contained by both.
+	\end{description}
+\end{defn}
+
+\begin{remark}
+	Highlight something important in the formal definitions or their use.
+\end{remark}
 
@@
 	\begin{tikzpicture}[scale = 0.3]
-		\draw (0,0) circle (3); 
+		\draw (0.3,0) circle (3); 
 		\draw (-0.5,-0.5) circle (1.5);
-		\node at (-4.7,0) {$A \subset B$};
+		\node at (-4.7,0) {$\circ A \subset A$};
 		\node at (-0.5,-0.5) {$A$};
-		\node at (1.3,1.3) {$B$};
+		\node at (1.3,1.3) {$\circ A$};
 	\end{tikzpicture} 
@@
 
-\begin{defn}[Set relationships]
-	Let $A$ and $B$ be two sets.
-\marginleft{Subset, superset: $\subseteq$, $\supseteq$} If all elements in $A$ are also contained in $B$, we say $A$ is a \textbf{subset} of $B$, noted $A \subseteq B$.
-In this case, we also say $B$ is a \textbf{superset} of $A$, noted $B \supseteq A$.
-Additionally, if $B$ contains elements that $A$ does not contain, we say $A$ is a \textbf{strict subset} of $B$, noted $A \subset B$, and $B$ is a \textbf{strict superset} of $A$, noted $B \supset A$.
-If the two sets contain exactly the same elements, then we say the sets are \textbf{equal}, noted $A = B$.
-If the two sets have no element in common they are said to be \textbf{disjoint}.
+\begin{defn}[Center - definition with a diagram]
+	The \textbf{center} of a set $A$ \marginleft{Set center: $\circ A$}, noted $\circ A$, is the subset of all blorped elements.
 \end{defn}
-
-\begin{defn}[Power set]
-	Given a set $A$, its \textbf{power set}, \marginleft{Power set:\\ $\mathcal{P}(A)$, $2^{A}$} noted $\mathcal{P}(A)$ or $2^{A}$, is the set of all subsets of $A$.
-That is, $\mathcal{P}(A) = \{B \, | \, B \subseteq A \}$.
-\end{defn}
-
-\marginright{
-	~\\
-	\def\setA{ (-1,0) circle (2) }
-	\def\setB{ (1,0) circle (2) }
-	\vspace{1pt}
-	\begin{tikzpicture}[scale = 0.3]
-		\filldraw[black!10] \setA;
-		\filldraw[black!10] \setB;
-		\draw \setA; 
-		\draw \setB;
-		\node at (-4.7,0) {$A\!\cup \! B$};
-	\end{tikzpicture} 
-	\vspace{1pt}
-	\begin{tikzpicture}[scale = 0.3]
-		\begin{scope}
-			\clip \setA;
-			\filldraw[black!10] \setB;
-		\end{scope}
-		\draw \setA; 
-		\draw \setB;
-		\node at (-4.7,0) {$A\!\cap \! B$};
-	\end{tikzpicture} 
-	\vspace{1pt}
-	\begin{tikzpicture}[scale = 0.3]
-		\filldraw[black!10] \setA;
-		\filldraw[white] \setB;
-		\draw \setA; 
-		\draw \setB;
-		\node at (-4.7,0) {$A\!\setminus \! B$};
-	\end{tikzpicture}
-	\def\setU{ (-3,-2) rectangle (3,2) }
-	\def\setA{ (0,0) circle (1.8) }
-	\begin{tikzpicture}[scale = 0.3] 
-		\filldraw[black!10] \setU;
-		\filldraw[white] \setA;
-		\draw \setU;
-		\draw \setA;
-		\node at (-5.3,0) {$A^{\complement}$};
-	\end{tikzpicture} 
-}
-
-\begin{defn}[Set operations]
-	We define \marginleft{Set operations: $\cup$, $\cap$, $\setminus$, $^{\complement}$} the following operations between two sets $A$ and $B$:
-	\begin{description}
-		\item[Union.]Noted $A \cup B$, the union of $A$ and $B$ is the set of all elements that are contained by $A$, $B$ or both.
-		\item[Intersection.] Noted $A \cap B$, the intersection of $A$ and $B$ is the set of all elements that are contained by both.
-		\item[Set difference (or relative complement).] Noted $A \setminus B$, $A$ minus $B$ is the set of all elements in $A$ that are not contained by $B$.
-		\item[Complement (or absolute complement).] Noted $A^{\complement}$, represents all elements that are not contained in $A$.
-Note that this operation is context specific: we need to know that, within a certain context, $U$ is the set of all elements under study; then the complement of $A \subseteq U$ is $A^{\complement} = U \setminus A$.
-	\end{description}
-
-	Let $\mathcal{A} \subseteq \mathcal{P}(A)$ be a set of subsets of $A$.
-The operations of union $\bigcup\limits_{A \in \mathcal{A}} A$ and intersection $\bigcap\limits_{A \in \mathcal{A}} A$ can be extend to span all elements.
-\end{defn}
-
-\begin{defn}[Ordered pair]
-	Given two \marginleft{Ordered pair: $(a, b)$} elements $a$ and $b$, the \textbf{ordered pair} $(a, b)$ specifies two objects in that order.
-That is $(a, b) \neq (b, a)$.
-This can be constructed setting $(a, b) = \{ \{a\}, \{a, b\}\}$.
-\end{defn}
-
-\marginright{~\\
-	\begin{tikzpicture}[scale = 0.3] 
-		\node at (-2.5,-4) {$A$};
-		\node at (3,2.3) {$B$};
-		\draw[rounded corners] (0,0.2) rectangle (6,1.5);
-		\draw[rounded corners] (-0.2,0) rectangle (-1.5,-8);
-		\draw[rounded corners] (0,0) rectangle (6,-8);
-
-		\def\circlepath{circle (0.45)}
-		\def\rectanglepath{ ++(-0.45,-0.45)  -- ++(0,0.9)  -- ++(0.9,0) -- ++(0,-0.9) -- ++(-0.9,0)}
-		\def\diamondpath{ ++(-0.5,0)  -- ++(0.5,0.5)  -- ++(0.5,-0.5) -- ++(-0.5,-0.5) -- ++(-0.5,0.5)}
-
-		% B elements
-		\draw (1,0.85) \circlepath;
-		\draw (3,0.85) \rectanglepath;
-		\draw (5,0.85) \diamondpath;
-
-		\definecolor{a}{rgb}{0.945, 0.415, 0.439}
-		\definecolor{b}{rgb}{0.694, 0.847, 0.466}
-		\definecolor{c}{rgb}{0.549, 0.862, 0.854}
-		\definecolor{d}{rgb}{0.301, 0.301, 0.301}
-
-		% A elements
-		\fill[a] (-0.85, -1) circle (0.5);
-		\fill[b] (-0.85, -3) circle (0.5);
-		\fill[c] (-0.85, -5) circle (0.5);
-		\fill[d] (-0.85, -7) circle (0.5);
-
-		% AxB elements
-		\filldraw[fill=a] (1, -1) \circlepath;
-		\filldraw[fill=b] (1, -3) \circlepath;
-		\filldraw[fill=c] (1, -5) \circlepath;
-		\filldraw[fill=d] (1, -7) \circlepath;
-		\filldraw[fill=a] (3, -1) \rectanglepath;
-		\filldraw[fill=b] (3, -3) \rectanglepath;
-		\filldraw[fill=c] (3, -5) \rectanglepath;
-		\filldraw[fill=d] (3, -7) \rectanglepath;
-		\filldraw[fill=a] (5, -1) \diamondpath;
-		\filldraw[fill=b] (5, -3) \diamondpath;
-		\filldraw[fill=c] (5, -5) \diamondpath;
-		\filldraw[fill=d] (5, -7) \diamondpath;
-
-		\node at (3,-9) {$A \times B$};
-	\end{tikzpicture} 
-}
-\begin{defn}[Cartesian product]
-	Given two sets \marginleft{Cartesian product: $\times$}  $A$ and $B$, the \textbf{Cartesian product} $A \times B = \{ (a,b) \, | \, \forall a \in A, b \in B \}$ is the set of all possible ordered pairs between the elements of $A$ and $B$.
-\end{defn}
-
-\begin{defn}[Disjoint union]
-	Let $A$ and $B$ be two sets.
-The disjoint union $A \sqcup B$ is the union where the elements of the sets are always treated as distinct.
-That is, given $A \sqcup B = \{0\} \times A \cup \{1\} \times B$.
-\end{defn}
-
-\begin{remark}
-	Ordered pairs and disjoint unions exemplify the effort in set theory to construct \emph{all object} starting from the primitives of set and set inclusion.
-\end{remark}
-
-\subsection{Relations and functions}
-
-\begin{remark}
-	Relations provide a common foundation for functions (i.e.
-the relation is the set of pairs $(x, f(x))$), equivalences and order (i.e.
-the relation is the set of pairs $(a, b)$ such that $a=b$ or $a\leq b$ respectively).
-This exemplifies the effort in set theory to find general structures.
-\end{remark}
-
-\marginright{~\\
-	\begin{tikzpicture}[scale = 0.3] 
-		\node at (-2.5,-4) {$A$};
-		\node at (3,2.3) {$B$};
-		\draw[rounded corners] (0,0.2) rectangle (6,1.5);
-		\draw[rounded corners] (-0.2,0) rectangle (-1.5,-8);
-		\draw[rounded corners] (0,0) rectangle (6,-8);
-
-		\def\circlepath{circle (0.45)}
-		\def\rectanglepath{ ++(-0.45,-0.45)  -- ++(0,0.9)  -- ++(0.9,0) -- ++(0,-0.9) -- ++(-0.9,0)}
-		\def\diamondpath{ ++(-0.5,0)  -- ++(0.5,0.5)  -- ++(0.5,-0.5) -- ++(-0.5,-0.5) -- ++(-0.5,0.5)}
-
-		% B elements
-		\node at (1,0.85) {\includegraphics[scale=0.25]{cat.png}};
-		\node at (3,0.85) {\includegraphics[scale=0.25]{dog.png}};
-		\node at (5,0.85) {\includegraphics[scale=0.25]{swan.png}};
-
-		\definecolor{a}{rgb}{1.0, 1.0, 1.0}
-		\definecolor{b}{rgb}{1, 0.509, 0}
-		\definecolor{c}{rgb}{0.45, 0.225, 0}
-		\definecolor{d}{rgb}{0, 0, 0}
-
-		% A elements
-		\filldraw[fill=a] (-0.85, -1) circle (0.5);
-		\fill[b] (-0.85, -3) circle (0.5);
-		\fill[c] (-0.85, -5) circle (0.5);
-		\fill[d] (-0.85, -7) circle (0.5);
-
-		% AxB elements
-		\node at (1, -1) {$\checkmark$};
-		\node at (1, -3) {$\checkmark$};
-		\node at (1, -5) {};
-		\node at (1, -7) {$\checkmark$};
-		\node at (3, -1) {$\checkmark$};
-		\node at (3, -3) {};
-		\node at (3, -5) {$\checkmark$};
-		\node at (3, -7) {$\checkmark$};
-		\node at (5, -1) {$\checkmark$};
-		\node at (5, -3) {};
-		\node at (5, -5) {};
-		\node at (5, -7) {$\checkmark$};
-
-		\node at (3,-9) {$R$};
-	\end{tikzpicture}
-
-	\begin{tikzpicture}[scale = 0.3] 
-		\node at (-2.5,-3) {$B$};
-		\node at (4,2.3) {$A$};
-		\draw[rounded corners] (0,0.2) rectangle (8,1.5);
-		\draw[rounded corners] (-0.2,0) rectangle (-1.5,-6);
-		\draw[rounded corners] (0,0) rectangle (8,-6);
-
-		\def\circlepath{circle (0.45)}
-		\def\rectanglepath{ ++(-0.45,-0.45)  -- ++(0,0.9)  -- ++(0.9,0) -- ++(0,-0.9) -- ++(-0.9,0)}
-		\def\diamondpath{ ++(-0.5,0)  -- ++(0.5,0.5)  -- ++(0.5,-0.5) -- ++(-0.5,-0.5) -- ++(-0.5,0.5)}
-
-		% B elements
-		\node at (-0.85, -1) {\includegraphics[scale=0.25]{cat.png}};
-		\node at (-0.85, -3) {\includegraphics[scale=0.25]{dog.png}};
-		\node at (-0.85, -5) {\includegraphics[scale=0.25]{swan.png}};
-
-		\definecolor{a}{rgb}{1.0, 1.0, 1.0}
-		\definecolor{b}{rgb}{1, 0.509, 0}
-		\definecolor{c}{rgb}{0.45, 0.225, 0}
-		\definecolor{d}{rgb}{0, 0, 0}
-
-		% A elements
-		\filldraw[fill=a] (1,0.85) circle (0.5);
-		\fill[b] (3,0.85) circle (0.5);
-		\fill[c] (5,0.85) circle (0.5);
-		\fill[d] (7,0.85) circle (0.5);
-
-		% AxB elements
-		\node at (1, -1) {$\checkmark$};
-		\node at (3, -1) {$\checkmark$};
-		\node at (5, -1) {};
-		\node at (7, -1) {$\checkmark$};
-		\node at (1, -3) {$\checkmark$};
-		\node at (3, -3) {};
-		\node at (5, -3) {$\checkmark$};
-		\node at (7, -3) {$\checkmark$};
-		\node at (1, -5) {$\checkmark$};
-		\node at (3, -5) {};
-		\node at (5, -5) {};
-		\node at (7, -5) {$\checkmark$};
-
-		\node at (4,-7) {$R^{-1}$};
-	\end{tikzpicture}
-
-}
-\begin{defn}[Binary relation]
-	Given a set $A$, \marginleft{Binary\\ relation: $aRb$} called \textbf{domain}, and a set $B$, called \textbf{codomain}, a \textbf{binary relation} is a set $R \subseteq A \times B$ of ordered pairs.
-We say $a$ is \textbf{$R$-related} to $b$, noted $aRb$ if $(a,b) \in R$.
-If $A=B$ (i.e.
-domain and codomain coincide) the relation is said \textbf{homogeneous}, \textbf{heterogeneous} if not.
-\end{defn}
-
-\begin{defn}[Image and preimage]
-	If $X \subseteq A$, \marginleft{Image and\\ preimage} the \textbf{image of $X$ under $R$} is the set of all elements in $B$ that are related to at least one element in $X$.
-The \textbf{image of $R$} is set of all elements in $B$ that are related to at least one element in $A$ (i.e.
-the image of the full set $A$).
-
-	If $Y \subseteq B$, the \textbf{preimage of $Y$ under $R$} is the set of all elements in $A$ that are related to at least one element in $Y$.
-The \textbf{preimage of $R$} is set of all elements in $A$ that are related to at least one element in $B$ (i.e.
-the preimage of the full set $B$).
-\end{defn}
-
-\begin{defn}[Converse relationship]
-	Given a relationship  \marginleft{Converse: $R^{-1}$} $R$ between $A$ and $B$, the \textbf{converse} or \textbf{inverse} relationship $R^{-1} \subseteq B \times A$ is the relationship where the order of the pairs (i.e.
-domain and codomain) is inverted.
-That is, $R^{-1} = \{ (b, a) \, | \, (a, b) \in R\}$.
-\end{defn}
-
-\begin{defn}[Relation properties]
-	Let $R$ \marginleft{Relation properties} be a binary relation between $A$ and $B$. We define the following properties:
-	\begin{description}
-		\item[Functional or right-unique.] An element $a \in A$ is related to at most one element $b \in B$. That is, if $aRb$ and $aRc$ then $b = c$.
-		\item[Injective or left-unique.] An element $b \in B$ is related to at most one element $a \in A$. That is, if $aRb$ and $cRb$ then $a = c$.
-		\item[Serial or left-total.] Every element $a \in A$ is related to at least one element $b \in B$.
-That is, for each $a \in A$ there exists at least one $b \in B$ such that $aRb$.
-In this case, the preimage of $R$ is the whole $A$.
-		\item[Surjective or right-total.] Every element $b \in B$ is related to at least one element $a \in B$.
-That is, for each $b \in B$ there exists at least one $a \in A$ such that $aRb$.
-In this case, the image of $R$ is the whole $B$.
-	\end{description}
-	From those basic properties, we define the following:
-	\begin{description}
-		\item[One-to-one.] Injective and functional (i.e. left-unique and right-unique).
-		\item[One-to-many.] Injective and not functional (i.e. left-unique and not right-unique).
-		\item[Many-to-one.] Not injective and functional (i.e. not left-unique and right-unique).
-		\item[Many-to-many.] Not injective nor functional (i.e. not left-unique nor right-unique).
-	\end{description}	
-\end{defn}
-
-\begin{defn}[Function]
-	A \textbf{partial function} \marginleft{(Partial) \\ functions: \\ $f : A \to B$} $f : A \to B$ is a binary relationship between $A$ and $B$ that is functional (i.e.
-right-unique).
-The \textbf{graph} of the function $G \subset A \times B$ is the function expressed as ordered pairs.
-A \textbf{total function}, or simply a \textbf{function}, is a partial function that is also serial (i.e.
-left-total, defined on the whole domain).
-A function is \textbf{bijective} if it is injective and surjective.
-
-	When the functional form $f(a)$ is known, this can be expressed as $a \mapsto f(a)$ (read ``$a$ maps to $f$ of $a$'').
-\end{defn}
-
-\begin{remark}
-	Note that the domain and the codomain can be any set, including scalar products.
-Therefore $f : \mathbb{R} \times \mathbb{R} \to \mathbb{R}$ for which $(x,y) \mapsto \sqrt{x^2 + y^2}$ gives us the Euclidean norm of a two dimensional vector.
-\end{remark}
-
-\begin{defn}[Identity function]
-	Given a set $A$, \marginleft{Identity: $\Id_A$} the identity function $\Id_A : A \to A$ is the function such that $\Id_A(a) = a$ for all $a \in A$.
-\end{defn}
-
-\begin{prop}[Inverse function]
-	Let $f : A \to B$ \marginleft{Inverse: $f^{-1}$} be a function.
-The corresponding converse relationship $f^{-1}$ is a function if and only if $f$ is bijective.
-In this case $f^{-1} : B \to A$ is also bijective and is called the \textbf{inverse} of $f$.
-\end{prop}
-
-\begin{defn}[Image/preimage functions]
-	Let $f : A \to B$ \marginleft{$f(X), f^{-1}(Y)$} be a function.
-This induces a relation $f : \mathcal{P}(A) \to \mathcal{P}(B)$ that associates each subset of $A$ to its image and a relation $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ that associates each subset of $B$ with its preimage.
-\end{defn}
-
-\begin{prop}[Properties of image/preimage functions]
-	The induced relations $f : \mathcal{P}(A) \to \mathcal{P}(B)$ and $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ are functions and satisfy the following properties:
-	\begin{enumerate}
-		\item $f(\bigcup\limits_{A \in \mathcal{A}} A) = \bigcup\limits_{A \in \mathcal{A}} f(A)$
-		\item $f^{-1}(\bigcup\limits_{A \in \mathcal{A}} A) = \bigcup\limits_{A \in \mathcal{A}} f^{-1}(A)$
-		\item $f^{-1}(\bigcap\limits_{A \in \mathcal{A}} A) = \bigcap\limits_{A \in \mathcal{A}} f^{-1}(A)$
-	\end{enumerate}
-\end{prop}
-
-\begin{remark}
-	With abuse of notation, $f$ and $f^{-1}$ indicate both the function between elements and sets.
-Note that $f^{-1} : \mathcal{P}(B) \to \mathcal{P}(A)$ always exists even if $f$ is not bijective.
-Also note that the image of the intersection is not in general the intersection of the image, and how maps that preserve set theoretic structures (e.g.
-topologies and $\sigma$-algebras) are defined based on reverse images.
-\end{remark}
-
-\begin{defn}[Fiber]
-	Given \marginleft{Fiber} a function $f : A \to B$, the A \textbf{fiber of $b \in B$ under $f$} is the subset of $A$ of where the function takes the value $b$.
-That is, the reverse image $f^{-1}(\{b\})$ of the singleton.
-Fibers over real functions are also called \textbf{level sets}.
-\end{defn}
-
-\begin{defn}[Restriction]
-	Let $f : A  \to B$ \marginleft{Restriction: $f|_C$} be a function and let $C \subseteq B$. Then the \textbf{restriction of $f$ over $C$} is the function $f|_C : C \to B$ such that $f|_C(c) = f(c)$.
-\end{defn}
-
-\begin{defn}[Composition]
-	Let $A$, $B$ and $C$ \marginleft{Function and relation composition: $R \circ S$, $f \circ g$} be three sets.
-Let $R$ be a binary relationship between $A$ and $B$ and $S$ be a binary relationship between $B$ and $C$.
-Then the \textbf{composition of $R$ and $S$} is the binary relationship $S \circ R \subseteq A \times C$ between $A$ and $C$ such that $(a, c) \in S \circ R$ if we can find $b \in B$ such that $aRb$ and $bSc$.
-\end{defn}
-
-\begin{remark}
-	The order of composition follows those of functions: if $f : A \to B$ and $g : B \to C$, composition leads to $c = g(f(a)) = (g \circ f)(a)$.
-\end{remark}
-
-\begin{prop}[Composition properties]
-	Relation composition \marginleft{Composition properties} follows the following properties:
-	\begin{itemize}
-		\item Composition is associative: $R \circ (T \circ S) = (R \circ T) \circ S$
-		\item Converse of composition: $(R \circ S)^{-1} = S^{-1} \circ R^{-1}$
-		\item Composition of functional/injective/serial/surjective relations is \\ functional/injective/serial/surjective.
-		\item If $f : A \to B$ and $g : B \to C$ are (partial) functions, then $g \circ f$ is a (partial) function.
-		\item Let $f : A \to B$ is a bijective function, then $f \circ f^{-1} = \Id_B$ and $f^{-1} \circ f = \Id_A$.
-	\end{itemize}
-\end{prop}
-
-\begin{defn}[Sets of functions]
-	The set \marginleft{Sets of functions: $B^A$} of all possible functions $f : A \to B$ is noted $B^A$.
-\end{defn}
-
-\begin{prop}
-	Let $2 = \{0, 1\}$ denote a set with two elements. Then $\mathcal{P}(A)$ and $2^A$ are in one-to-one correspondence.
-\end{prop}
-
-\begin{remark}
-	The notation for function sets is due to the cardinality.
-If $A$ has $n$ elements and $|B|$ has $m$ elements, each function is given by choosing an element of $B$ for each element of $A$, that is $m$ choices $n$ times, that is $m^n$.
-The power set is noted $2^A$ because each subset can be identified by a function $U : A \to \{0 , 1\}$, where $f(a)=0$ corresponds to $a \notin U$ and $f(a)=0$ corresponds to $b \in U$.
-\end{remark}
-
-\begin{defn}[Set closure under operations]
-	Let $f : A_1 \times A_2 \times ...
-\times A_n \to B$ be an $n$-ary function.
-We say a set \textbf{$A$ is closed under $f$} if whenever the function is applied to elements of $A$ returns an element of $A$.
-That is, $A \subseteq A_i$ for all $1\leq i \leq n$, $A \subset B$ and $f(A, A, ..., A) \subseteq A$.
-\end{defn}
-
-\begin{remark}
-	For example, integers are closed under addition, subtraction and multiplication but are not closed under division.
-Most mathematical structures (groups, topologies, $\sigma$-algebra, vector spaces, ...) define sets closed under some operations.
-\end{remark}
-
-\subsection{Indexed families and sequences}
-
-\begin{defn}[Families]
-	An \textbf{indexed family}, \marginleft{Families, sequences: $\{x_i\}_{i \in I}$, $\{x_i\}_{i=1}^{\infty}$} or simply a \textbf{family}, $\{x_i\}_{i \in I}$ is a collection of elements identified by an \textbf{index set} $I$.
-More formally, an indexed family is a tuple $\langle X, I, x \rangle$ where $X$ and $I$ are sets and $x: I \to X$ is a function such that $i \mapsto x_i$.
-A \textbf{sequence} is a family for which the index set is a continuous interval of natural numbers.
-A \textbf{finite sequence} of $n$ elements is noted $\{x_i\}_{i=1}^{n}$ while an \textbf{infinite sequence} is noted $\{x_i\}_{i=1}^{\infty}$.
-\end{defn}
-
-\begin{defn}[Operations on families of sets]
-	Let $\{A_i\}_{i \in I}$ \marginleft{$\bigcap\limits_{i \in I} A_i $, $\bigcup\limits_{i \in I} A_i$} be a family of sets, the operations of intersection $\bigcap\limits_{i \in I} A_i $ and union $\bigcup\limits_{i \in I} A_i$ can be extended to span all members.
-\end{defn}
-
-\begin{defn}[Set-theoretic limit]
-	Let $\{A_i\}_{i=0}^{\infty}$ \marginleft{Set limit: \\ $\lim\limits_{i \to \infty} A_i = A$} be a sequence of sets.
-We define the \textbf{limit infimum} as $\liminf\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1}\bigcap\limits_{j \geq i} A_j$ and the \textbf{limit supremum} as $\limsup\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1}\bigcup\limits_{j \geq i} A_j$.
-If the two are equal to the same set then we say the sequence \textbf{converges}, the \textbf{set-theoretic limit} $\lim\limits_{i \to \infty} A_i = A$ exists and it is equal to that set.
-\end{defn}
-
-\marginright{~\\
-	\def\setA{ (-1,0) circle (2) }
-	\def\setB{ (1,0) circle (2) }
-	\begin{tikzpicture}[scale = 0.3]
-		\draw (0,0) ellipse (5 and 4); 
-		\draw (0.5,-0.7) ellipse (3.5 and 2.7);
-		\draw (1.05,-1.3) ellipse (2.2 and 1.6);
-		\draw (1.25,-1.45) ellipse (1.6 and 1.05);
-		\draw (1.3,-1.6) ellipse (1.05 and 0.55);
-		\node at (0,-5) {$A_i \supset A_{i+1}$};
-		\node at (-2.6, 2.1) {$A_1$};
-		\node at (-1.05,0.55) {$A_2$};
-		\node at (1.3,-1.6) {$...$};
-	\end{tikzpicture} 
-}
-\begin{defn}[Monotone sequence]
-	A \textbf{monotone sequence} \marginleft{Monotone sequence: \\ $A_1 \subseteq A_2 \subseteq ...$\\ $A_1 \supseteq A_2 \supseteq ...$} is a sequence of sets $\{A_i\}_{i=0}^{\infty}$ where each set is a subset (or superset) of the preceding.
-More specifically, we distinguish four cases:
-	\begin{description}
-		\item[Increasing] when $A_i \subseteq A_{i+1}$ for all $i>0$
-		\item[Decreasing] when $A_i \supseteq A_{i+1}$ for all $i>0$
-		\item[Strictly increasing] when $A_i \subset A_{i+1}$ for all $i>0$
-		\item[Strictly decreasing] when $A_i \supset A_{i+1}$ for all $i>0$
-	\end{description}
-\end{defn}
-
-\begin{prop}[Monotone convergence]
-	All monotone \marginleft{Monotone convergence} sequences converge.
-For increasing sequences the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1} A_i$, while for decreasing sequences  the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1} A_i$.
-\end{prop}
-
-\subsection{Homogeneous relations, equivalences and orders}
-
-\begin{remark}
-	We will cover the notions of orders that are necessary for the definition of ordinals and cardinal numbers. Orders will be treated more in detail in a Bare Minimum dedicated to order theory.
-\end{remark}
-
-\begin{defn}[Homogeneous relation properties]
-	Let $R \subseteq A \times A$ \marginleft{Reflexivity, symmetry, transitivity, ...} be a homogeneous binary relation between $A$ and itself. We define the following properties:
-	\begin{description}
-		\item[Reflexive.] Every element $a \in A$ is related to itself. That is, $aRa$ for all $a \in A$.
-		\item[Irreflexive.] No element $a \in A$ is related to itself. That is, $aRa$ for no $a \in A$.
-		\item[Symmetric.] If $a$ is related to $b$, then $b$ is related to $a$. That is, if $aRb$ then $bRa$.
-		\item[Antisymmetric.] If $a$ is related to $b$ and $b$ is related to $a$, then $a$ and $b$ are the same element. That is, if $aRb$ and $aRb$ then $a=b$.
-		\item[Transitive.] If $a$ is related to $b$ and $b$ is related to $c$, then $a$ is related to $c$. That is, if $aRb$ and $bRc$ then $aRc$.
-		\item[Total.] Given two distinct elements, one is related to the other.That is, given $a \neq b$, either $aRb$ or $bRa$. (This has no relation to left/right-total)
-	\end{description}
-
-\end{defn}
-
-\begin{defn}[Orders and equivalence]
-	A \textbf{partial order} (noted $\leq$, $\lesssim$, $\preceq$) \marginleft{Order, equivalence: $\leq$, $<$, $\equiv$} is a homogeneous binary relationship that is reflexive, antisymmetric and transitive.
-A \textbf{linear order} or \textbf{total order} is an order that is total.
-An order is \textbf{strict} (noted $<$, $\prec$) if it is irreflexive instead of reflexive.
-
-	An \textbf{equivalence relation} (noted $\equiv$, $\sim$) is a homogeneous binary relationship that is reflexive, symmetric and transitive.
-\end{defn}
-
-\begin{defn}[Partitions]
-	A \textbf{partition} \marginleft{Partitions} of a set $A$ is a collections $P \subset \mathcal{P}(A)$ of disjoint subsets that cover all of $A$. That is:
-	\begin{enumerate}
-		\item $\bigcup\limits_{B \in P} B = A$
-		\item $B \cap C = \emptyset$ for all $B, C \in P$ where $B \neq C$
-	\end{enumerate}
-\end{defn}
-
-\begin{defn}[Equivalence classes and quotient set]
-	Let $A$ \marginleft{Eq.
-classes and quotient set: \\ $[a]_{\sim}, a_{/\sim}, A_{/\sim}$} be a set and $\sim$ an equivalence relation on $A$.
-Let $a \in A$.
-The \textbf{equivalence class of $a$ by $\sim$} is the set $[a]_{\sim} = a_{/\sim} = \{b \in A \, | \, b \sim a \}$ of all the elements of $A$ equivalent to $a$.
-The \textbf{quotient set of $A$ by $\sim$} $A_{/\sim} = \{ a_{/\sim} \, | \, a \in A \}$  is the set of all of the equivalences classes by $\sim$.
-\end{defn}
-
-\begin{prop}[Quotient sets are partitions]
-	Every quotient set is a partition, and every partition is a quotient set.
-That is, let $A$ be a set and $\sim$ an equivalence class, then $A_{/\sim}$ is a partition.
-Conversely, let $P$ be a partition of $A$, then $\sim \; = \{ (a,b) \, | \, a, b \in X \text{ for some } X \in P \} $ is an equivalence relation and $A_{/\sim} = P$.
-\end{prop}
-
-\begin{defn}
-	Given a function $f : A \to B$ the \textbf{equivalence relation determined by $f$} is the relation $R_f = \{ (a_1, a_2) \, | \, f(a_1) = f(a_2) \}$.
-\end{defn}
-
-\begin{prop}[Function canonical decomposition]
-	Every function $f : A \to B$ can be written as the composition $f = t \circ s \circ r$ of the following three functions:
-	\begin{description}
-		\item[$r : A \to A_{/R_f}$] defined as $a \mapsto [a]_{R_f}$; this function is surjective
-		\item[$s : A_{/R_f} \to f(A)$] defined as $[a]_{R_f} \mapsto f(a)$; this function is bijective
-		\item[$t : f(A) \to B$] defined as $t(b) = b$; this function is injective.
-	\end{description}
-\end{prop}
-
-\begin{remark}
-	All the above concepts are of general use in all areas of mathematics. Note that they are all constructed on two primitives: the notion of sets and of set inclusion.
-\end{remark}
-
-\section{Axiomatic set theory}
-
-Axiomatic set theory was introduced to eliminate some technical problems that would make naive set theory inconsistent.
-There is no single approach and there is no single set of axioms.
-For example, constructive set theory, as all constructive mathematics, rejects ``nonconstructive proofs'' in which the existence of an object is proved with no explicit algorithm as to how to construct it.
-This will lead to different choices than the ones outlined here.
-Much of the work on the foundations of mathematics in set theory is to clarify the relationship between the axioms: which ones are required to get certain results, which ones are independent and which ones are inconsistent.
-
-This is an important insight for those working on the Assumptions of Physics: the foundations of mathematics are not absolute.
-We have to be conscious of the consequences of those choice and understand whether they are suitable for the foundations of physics.
-
-\subsection{Paradoxes}
-
-While they are generally called paradoxes, the following are really contradictions and that is why axiomatic set theory needs to avoid them.
-The first type of paradoxes to naive set theory arise from the use of natural language and the meaning of the words, typically when this is self-referential.
-
-\textbf{Berry's paradox}.
-Consider the expression ``the smallest positive integer not definable in under sixty letters.'' Given that there are only twenty-six letters in the English language, there are finitely many phrases under sixty letters and therefore only finitely many integers can be defined in under sixty letters.
-Therefore there are integers not definable under sixty letters and, since the integers are ordered, there will be a least one.
-However, the above expression contains fifty-seven letters.
-We have a contradiction.
-
-These type of paradoxes are solved by restricting to symbolic expressions of a formal language, such as predicate logic.
-The consequence is that, at the foundational level, mathematics does not keep track of what the symbols mean.
-
-The first type of paradoxes are purely logical, so they can be constructed in the formal system.
-
-\textbf{Russell's paradox}.
-Let $R$ be the set of all sets that contain themselves.
-That is, $R=\{ x \, | \, x \notin x\}$.
-Suppose $R \notin R$ then, by definition, $R \in R$.
-Suppose $R \in R$ then, by definition, $R \notin R$.
-We have a contradiction.
-
-These are solved by carefully specifying the rules that are allowed for the creation of sets, by picking the ``right'' axioms. Part of the solution revolves around the following concepts.
-
-%https://en.wikipedia.org/wiki/Zermelo%E2%80%93Fraenkel_set_theory
-
-%https://en.wikipedia.org/wiki/Von_Neumann–Bernays–Gödel_set_theory
-
-\subsection{Classes, sets and elements}
-
-A \textbf{class} \marginright{ ~\\
-		\begin{tikzpicture}[scale = 0.3]
-		\node [draw, inner sep=6, rectangle] (class) {\small Class};
-		\node [below=.3 of class] (middle) {};
-		\node [below=1.3 of class] (lower) {};
-		\node [draw, inner sep=6, rectangle, align=center, left=0 of lower] (pclass) {\small Proper \\ class};
-		\node [draw, inner sep=6, rectangle, right=.2 of lower] (set) {\small Set};
-		\draw[->] (pclass.north) -| ++(0,.8) -|  (class.south);	
-		\draw[->] (set.north) -| ++(0,1.5) -|  (class.south);	
-		\draw[->] (class.east) -| ++(3,0) |-  (set.east);	
-		 %++(0,-1) -|
-		\node [left=-0.1 of middle] (isa) {\small is a};
-		\node at (4.1,.6) (contains) {\small contains};
-	\end{tikzpicture} 
-}  is any collection, therefore \textbf{sets are classes}.
-Some classes are not sets.
-These are called \textbf{proper classes} and are exactly the collections that cannot be elements of a class.
-Therefore, sets are classes that can be contained by other classes, while proper classes cannot.
-The difference between sets and proper classes is not in what they contain but in whether they can be contained.
-Proper classes are ``too large'' to be sets in the following sense: given a class $A$, if we can find a set $B$ that contains all elements of $A$, then $A$ will be a set as well.
-Therefore proper classes can never be subclasses of sets, they are bigger than all sets.
-The class of all sets, for example, is a proper class.
-So is the class of all cardinal numbers.
-
-Note that axiomatic set theory does not distinguish between objects that can or cannot contain other objects.
-All elements are classes, all element can, in principle, contain something.
-In fact, since equality of two objects coincides on whether they have the same elements, two objects must contain something different to be different.
-Numbers will be constructed as sets starting from the empty set.
-As a consequence, the proposition $3 \in 4$ is valid, and happens to be true if $3$ and $4$ are naturals, though it is not true if $3$ and $4$ are reals.
-From the perspective of the Assumptions of Physics, this is not great as the framework allows propositions that are not physically meaningful.
-Radical different approaches to the foundations of mathematics (e.g.
-through category theory) do not exhibit this particular problem, though they exhibit others (e.g.
-identity of an object can be defined only up to its relationships with other objects).
-
-\subsection{Axiomatization}
-
-One strategy for the axiomatization of set theory, refered to as Zermelo–Fraenkel (ZFC), only defines sets.
-It avoids paradoxes by restricting set construction to subsets only.
-Proper classes are not objects that live in the framework.
-
-Another strategy, refered to as von Neumann–Bernays–Gödel (NBG), starts by defining classes and then defines sets as those classes that can be elements of classes.
-It avoids paradoxes by restricting set construction on predicates over sets.
-Since a proper class cannot be an element, if the definition creates a contradiction for an item $x$ (i.e.
-we have both $x \in S$ and $x \notin S$) than it must be that the element $x$ is a proper class.
-
-The two approaches are proved to be equivalent.
-
-\textbf{Axioms of set theory}. Here \marginleft{Set theory axioms} are the axioms of a particular formulation of axiomatic set theory taken from \cite{pinter2014book} that follows NBG:
-\begin{enumerate}
-	\item Axiom of extent. If two classes have the same elements, then they are identical.
-	\item Axiom of class construction.
-Let $P(x)$ be a predicate expressed in terms of the symbols $\in, \OR, \AND, \NOT, \Rightarrow, \exists, \forall$, brackets, variables $x, y, z, A, B, ...$ Then there exists a class $C$ that consists of all the elements (i.e.
-sets) x which satisfy $P(x)$.
-	\item Every subclass of a set is a set.
-	\item $\emptyset$ is a set.
-	\item If $a$ and $b$ are sets, then $\{a, b\}$ is a set.
-	\item If $\mathcal{A}$ is a set of sets, then, $\bigcup\limits_{A \in \mathcal{A}} A$ is a set.
-	\item If $A$ is a set, the power set $\mathcal{P}(A)$ is a set.
-	\item Axiom of foundation. If $A \neq \emptyset$ is a set, there exists an element $a \in A$ such that $a \cap A = \emptyset$.
-	\item Axiom of replacement. If $A$ is a set and $f : A \to B$ is a surjective function, then $B$ is a set.
-	\item Axiom of choice. Every set has a choice function. That is, given a set $A$, there exists a function $f : \mathcal{P}(A) \setminus \{ \emptyset \} \to A$ such that $f(B) \in B$.
-	\item Axiom of infinity. There exists a successor set (as defined in \ref{defn_successor_set}).
-\end{enumerate}
-
-\begin{remark}
-	Axiom 1 serves to define class equality.
-Axiom 2 defines how to define classes based on predicates.
-Axioms 3-7 tells us that some classes, constructable by axiom 2, are sets.
-Axiom 8 guarantees that we cannot have an infinite sequences $\{ A_i \}_{i=1}^\infty$ such that $A_{i+1} \in A_i$.
-In particular, a set cannot be an element of itself.
-Axiom 9 tells us that any set that is ``smaller'' than a set (i.e.
-we can map another set onto it) is a set.
-Axiom 10 guarantees that we can always pick an element from a set, though it does not tell us how.
-Axiom 11 tells us that there are sets with infinite elements.
-Once the natural numbers are assumed to exist, higher cardinality of infinity can be constructed through, for example, the power set.
-\end{remark}
-
-The axiom of choice has been the focus of particular investigation as it has been regarded as controversial for some of its consequences.
-While most mainstream mathematicians accept it, constructive mathematics rejects the axiom of choice as it does not lead to constructive proofs: it claims some objects exist even though they may not even be expressible or computable.
-It also implies the law of excluded middle (i.e.
-a proposition is either true or not), which constructive mathematics rejects.
-What does or does not require the axiom of choice is therefore ample subject of study.
-
-\begin{prop}
-	The following statements are equivalent:
-	\begin{itemize}
-		\item Axiom of choice: every set has a choice function
-		\item Well-ordering theorem: every set can be well ordered
-		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other
-		\item Zorn's lemma: every non-empty poset (i.e. partially ordered set) for which every chain (i.e. totally ordered subset) has an upper bound contains at least one maximal element
-		\item Every vector space has a basis
-	\end{itemize}
-\end{prop}
-
-\section{Numeric constructions}
-In set theory, numbers are defined to be sets with particular properties.
-These constructions are instructive to gain insights in the foundations of mathematics.
-Note that, for the Assumptions of Physics, we found more productive to understand numbers as order-theoretic concept (i.e.
-the integer numbers can be thought as a linearly ordered set where each element has an immediate successor and an immediate predecessor).
-
-\subsection{Standard numbers}
-
-\begin{defn}\label{defn_successor_set}
-	Let $A$ be a set.
-We define the \textbf{successor} $A^+ = A \cup \{A\}$.
-A set $S$ is called a \textbf{successor set} if it contains the empty set (i.e.
-$\emptyset \in S$ and if contains the successor of every element (i.e.
-if $X \in S$ then $X^+ \in S$).
-The set of \textbf{natural numbers} $\mathbb{N}$ is the intersection of all successor sets.
-The set of \textbf{positive naturals} is $\mathbb{N}^+ = \mathbb{N} \setminus \{ \emptyset \}$.
-\end{defn}
-\begin{remark}
-	The previous set exists only because one of the axioms tells us it does.
-The opposite axiom can be consistently chosen.
-The natural numbers so constructed are known as von Neumann ordinals.
-They are as follows:
-	\begin{itemize}
-		\item $0 = \{ \} = \emptyset$
-		\item $1 = \{ 0 \} = \{ \emptyset \}$
-		\item $2 = \{0, 1\} = \{ \emptyset, \{ \emptyset\} \}$
-		\item $3 = \{0, 1, 2\} = \{ \emptyset, \{ \emptyset\}, \{ \emptyset, \{ \emptyset\} \} \}$
-		\item $4 = \{0, 1, 2, 3\} = \{ \emptyset, \{ \emptyset\}, \{ \emptyset, \{ \emptyset\} \}, \{ \emptyset, \{ \emptyset\}, \{ \emptyset, \{ \emptyset\} \} \} \}$
-		\item ...
-	\end{itemize}
-	It can be shown that they obey Peano's axioms, another common way to axiomatize the natural numbers and their arithmetic.
-\end{remark}
-
-\begin{defn}
-	The \textbf{integers} can be defined from the natural number as an ordered pair of sign and value. That is, $\mathbb{Z} = \{ 0 \} \times \mathbb{N} \cup \{ 1 \} \times \mathbb{N}^+$.
-\end{defn}
-
-\begin{defn}
-	The \textbf{rationals} can be defined from the integers number as the equivalence class that leads to the same simplification.
-That is, let $Q = \mathbb{Z} \times \mathbb{Z}$ be all pairs of integers.
-Let $\sim \, \subset Q \times Q$ be an equivalence relation such that $(a_1, b_1) \sim (a_2, b_2)$ if and only if $a_1 b_2 = a_2 b_1$ (i.e.
-the pairs are proportional).
-The set of rationals is $\mathbb{Q} = Q_{/\sim}$.
-Given two integers $a, b \in \mathbb{Z}$, the rational number $\frac{a}{b}$ is the equivalence class $[(a,b)]_{\sim}$.
-\end{defn}
-
-\begin{defn}
-	The \textbf{reals} can be defined from the rationals as ``cuts'' that divide the rational in two contiguous sets.
-That is, let a real number be a partition $(A, B)$ such that $A \cup B = \mathbb{Q}$, $A \neq \emptyset$, $B \neq \emptyset$ and $a < b$ for all $a \in A$ and $b \in B$.
-The set of all reals $\mathbb{R}$ is the set of all such partitions.
-\end{defn}
-
-\begin{remark}
-	Operations such as additions, multiplication, ... can be defined appropriately as functions of the respective sets.
-\end{remark}
-
-\subsection{Ordinal numbers}
-
-To extend natural numbers beyond finite numbers we need to make a distinction.
-Cardinal numbers (i.e.
-one, two, three, ...) identify the number of elements in a set.
-Ordinal numbers (i.e.
-first, second, third) identify a position within a sequence.
-Cardinals and ordinals are in one to one correspondence only for finite values.
-For infinite values, cardinals are a subset of the ordinals.
-
-\begin{remark}
-	To define ordinals, we extend the notion of sequences to arbitrary levels of infinity through well-ordered sets.
-\end{remark}
-
-\begin{defn}[Poset]
-	A \textbf{partially ordered set}, \marginleft{Poset: $(A, \leq_A)$} or \textbf{poset}, is a set with an associated partial order.
-That is, a tuple $(A, \leq_A)$ where $A$ is a set and $\leq_A$ is a partial order over $A$.
-A \textbf{linearly ordered set} is a poset whose order is linear.
-\end{defn}
-
-\begin{defn}[Initial segment]
-	The \textbf{initial segment} \marginleft{Initial segment} of a poset $(A, \leq_A)$ determined by $a \in A$ is the set of all elements before $a$. That is, $S_a=\{ x \in A \, | \, x < a \}$.
-\end{defn}
-
-\begin{defn}[Order morphisms]
-	Let $(A, \leq_A)$ \marginleft{Order homo/ \\ isomorphism} and $(B, \leq_B)$ be two posets.
-An \textbf{order homomorphism} is a function $f : A \to B$ that preserves the ordering.
-That is, $a_1 \leq_A a_2$ implies $f(a_1) \leq_B f(a_2)$.
-An \textbf{order isomorphism} is a bijective order homomorphism whose inverse is also an order homomorphism (i.e.
-the posets have identical order).
-That is, $b_1 \leq_B b_2$ implies $f^{-1}(b_1) \leq_A f^{-1}(b_2)$.
-\end{defn}
-
-\begin{defn}[Well-order]
-	A \textbf{well-ordered} \marginleft{Well-order} set is a linearly ordered set $(A, \leq_A)$ where each subset has a least (i.e.
-first) element.
-That is, for every $B \subseteq A$, there exists an $a \in B$ such that $a \leq_A b$ for all $b \in B$.
-Let $A$ and $B$ be two well ordered set, we say $A$ \textbf{has lower ordinality than} $B$, noted $A \ordleq B$, if $A$ is orderwise identical to the beginning of $B$.
-That is, $A$ is order isomorphic to an initial segment of $B$.
-They have \textbf{the same ordinality}, noted $A \ordeq B$, if they have the same order.
-That is, they are order isomorphic.
-\end{defn}
-
-\begin{prop}[Ordinality]
-	If $A$ and $B$ are two well ordered sets, one of them has lower cardinality than the other.
-That is, either $A \ordleq B$ or $B \ordleq A$.
-If $\mathcal{A}$ is a collection of well-ordered sets, then one element has least ordinality.
-\end{prop}
-
-\begin{remark}
-	Simple examples of well-ordered sets:
-	\begin{itemize}
-		\item a finite sequence: $(a_1, a_2, ..., a_n)$
-		\item an infinite sequence: $(a_1, a_2, ...)$
-		\item an infinite sequence with elements afterwards: $(a_1, a_2, ..., b_1, b_2, ... , b_6)$
-		\item three infinite sequences, one after the other: $(a_1, a_2, ..., b_1, b_2, ... , c_1, c_2, ...)$
-	\end{itemize}
-	In a well-ordered set, every element always has an immediate successor, but not necessarily an immediate predecessor (e.g. $b_1$ in the third example).
-\end{remark}
-
-\begin{defn}[Ordinals]
-	An \textbf{ordinal} \marginleft{Ordinals: \ordinals} is a set $A$ that includes the previous and only the previous ordinals (this makes $a \in A$ mean $a \ordless A$).
-That is, for all elements $a, b, c \in A$ (i.e.
-previous ordinals of $A$):
-	\begin{enumerate}
-		\item $a \notin a$ (each previous ordinal is not before itself)
-		\item $a \in b$ implies $b \notin a$ (if a previous ordinal is before another, then the latter is not before the former)
-		\item $a \in b$ and $b \in c$ implies $a \in c$ (previous ordinals are transitive)
-		\item $x \in a$ implies $x \in A$ (an ordinal before a previous ordinal is a previous ordinal)
-	\end{enumerate}
-	We note $\ordinals$ as the class of all ordinals.
-\end{defn}
-
-\begin{prop}[Natural numbers are ordinals]
-	The natural numbers are ordinals. That is, $\mathbb{N} \subset \ordinals$.
-\end{prop}
-
-\begin{remark}
-	Ordinal numbers so defined generalize the construction of von Neumann ordinals (i.e. a natural number is a set that includes all previous natural numbers) to all possible well-orders.
-\end{remark}
-
-\begin{prop}[Ordinals are well-ordered]
-	All ordinals are well-ordered sets.
-That is, for each $A \in \ordinals$, $(A, \leq_A)$ is a well-ordered set where $a \leq b$ if either $a \in b$ or $a = b$.
-Moreover, the class $\ordinals$ is well ordered.
-That is, given a set of ordinals, one has lower ordinality than all others.
-\end{prop}
-
-\begin{prop}[Well-orders are ordinals]
-	For every \marginleft{$\langle A \rangle$, $\omega$} well-ordered set $A$ there is a unique ordinal that fully characterizes its order.
-That is, there is a unique $\langle A \rangle \in \cardinals$ such that $\langle A \rangle \ordeq A$.
-We say $\langle A \rangle$ is the \textbf{ordinality of $A$}.
-The ordinality of an infinite sequence is denote by $\omega$.
-That is, $\omega = (0, 1, 2, ...) = \langle \{x_i\}_{i=1}^\infty \rangle$ for all $\{x_i\}_{i=1}^\infty$.
-\end{prop}
-
-\begin{remark}
-	As an example, consider the von Neumann ordinal $3 = (0, 1, 2)$ as a well-ordered set.
-Similarly, the ordinality of the well-ordered set $(a_1, a_2, a_3)$ is $3 = (0, 1, 2)$ as they share the same ordering.
-The definitions are such that these ideas extend to infinite sets.
-\end{remark}
-
-\begin{defn}[Ordinal addition]
-	The addition \marginleft{Ordinal addition: $\langle A \rangle + \langle B \rangle$} of two ordinals is defined by concatenating the first order with the second.
-Let $(A, \leq_A)$ and $(B, \leq_B)$ be two well ordered sets.
-Let $(A \sqcup B, \leq)$ be the poset where:
-	\begin{itemize}
-		\item $(0, a_1) \leq (0, a_2)$ if $a_1 \leq_A a_2$
-		\item $(1, b_2) \leq (1, b_2)$ if $b_1 \leq_B b_2$
-		\item $(0, a) \leq (1, b)$ for all $a \in A$ and $b \in B$.  	
-	\end{itemize}
-	Then $(A \sqcup B, \leq)$ is well-ordered and the ordinal $\langle A \sqcup B \rangle$ is fully determined by $\langle A \rangle$ and $\langle B \rangle$.
-Therefore we can define the \textbf{ordinal addition} $\langle A \rangle + \langle B \rangle =  \langle A \sqcup B \rangle$.
-\end{defn}
-
-%\begin{prop}[Ordinal multiplication]The addition has the following properties:
-%	\begin{enumerate}
-%		\item associative: $\langle A \rangle + ( \langle B \rangle + \langle C \rangle) = (\langle A \rangle + \langle B \rangle) + \langle C \rangle$
-%		\item $\langle 0 \rangle = \langle \emptyset \rangle$ is the identity: $\langle A \rangle + \langle 0 \rangle = \langle 0 \rangle + \langle A \rangle = \langle A \rangle$
-%		\item not commutative in general: $\langle A \rangle + \langle B \rangle$ not necessarily the same as $\langle B \rangle + \langle A \rangle$
-%		\item reduces to arithmetic addition on the natural numbers
-%	\end{enumerate}
-%\end{prop}
-
-\begin{remark}
-	Addition in non-commutative (i.e.
-$\langle A \rangle + \langle B \rangle \neq \langle B \rangle + \langle A \rangle$) for infinite ordinals.
-For example $(a_1, a_2, ...) \sqcup (b_1, b_2, b_3) = (a_1, a_2, ..., b_1, b_2, b_3)$ is not the same as $(b_1, b_2, b_3) \sqcup (a_1, a_2, ...) = (b_1, b_2, b_3, a_1, a_2, ...) \ordeq (c_1, c_2, ...)$.
-Therefore $\omega + 3 \neq 3 + \omega = \omega$.
-\end{remark}
-
-\begin{defn}[Ordinal multiplication]
-	The multiplication \marginleft{Ordinal multiplication: $\langle A \rangle \langle B \rangle$} of two ordinals is defined by ordering their Cartesian product.
-Let $(A, \leq_A)$  and $(B, \leq_B)$ be two well ordered sets.
-Let $(A \times B, \leq)$ be the poset where:
-	\begin{itemize}
-		\item $(a_1, b_1) \leq (a_2, b_2)$ if $b_1 \leq_B b_2$
-		\item $(a_1, b) \leq (a_2, b)$ if $a_1 \leq_A a_2$
-	\end{itemize}
-	Then $(A \times B, \leq)$ is well-ordered and the ordinal $\langle A \times B \rangle$ is fully determined by $\langle A \rangle$ and $\langle B \rangle$.
-Therefore we can define the \textbf{ordinal multiplication} $\langle A \rangle \langle B \rangle =  \langle A \times B \rangle$.
-\end{defn}
-
-%\begin{prop}[Ordinal multiplication]
-%	The multiplication has the following properties:
-%	\begin{enumerate}
-%		\item associative: $\langle A \rangle ( \langle B \rangle \langle C \rangle) = (\langle A \rangle \langle B \rangle) \langle C \rangle$
-%		\item $\langle 1 \rangle = \langle \{ \emptyset \} \rangle$ is the identity: $\langle A \rangle \langle 1 \rangle  = \langle 1 \rangle \langle A \rangle = \langle A \rangle$
-%		\item right distributive over addition: $\langle A \rangle ( \langle B \rangle + \langle C \rangle) = \langle A \rangle \langle B \rangle + \langle A \rangle \langle C \rangle$
-%		\item $\langle A \rangle \langle 0 \rangle  = \langle 0 \rangle \langle A \rangle = \langle 0 \rangle$
-%		\item not commutative in general: $\langle A \rangle + \langle B \rangle$ not necessarily the same as $\langle B \rangle + \langle A \rangle$
-%		\item reduces to arithmetic multiplication on the natural numbers
-%	\end{enumerate}
-% The multiplication is associative and $\langle \emptyset \rangle$ is the identity (i.e.
-$\langle A \rangle + \langle \emptyset \rangle = \langle \emptyset \rangle + \langle A \rangle = \langle A \rangle$).
-The additions is not, in general, commutative.
-%\end{prop}
-
-\begin{remark}
-	Muliplication is non-commutative (i.e.
-$\langle A \rangle \langle B \rangle \neq \langle B \rangle \langle A \rangle$) for infinite ordinals.
-For example $(a_1, a_2, ...) \times (b_1, b_2) = ((a_1, b_1), (a_2, b_1), ..., (a_1, b_2), (a_2, b_2), \\...)$ is not the same as $(b_1, b_2) \times (a_1, a_2, ..., a_n, ...) = ((b_1, a_1), (b_2, a_1), (b_1, a_2), \\ (b_2, a_2), ...) \ordeq (c_1, c_2, ...)$.
-Therefore $\omega 2 \neq 2 \omega = \omega$.
-
-	With addition and multiplication, we can get a sense of how the ordinals work.
-$\omega$ represents a countable sequence (i.e.
-$\omega = (0, 1, 2, ...) = (\mathbb{N}, \leq)$).
-$\omega +3$ is a countable sequence followed by three elements.
-$\omega + \omega  = \omega2$ represents a countable sequence followed by a countable sequence.
-$\omega6$ represents a sequence of 6 countable sequences.
-$\omega\omega = \omega^2$ represents a countable sequence of countable sequences.
-$\omega^3$ represents a countable sequence of countable sequences of countable sequences.
-$\omega^\omega$ extends the recursion infinitely many times and so on.
-There is no upper limit to the ordinals.
-
-	Adding an ordinal of a lower power to the left does not change the ordinal.
-Adding finitely many elements at the beginning of an infinite sequence, still gives us an infinite sequence (i.e.
-$n + \omega = \omega$).
-In the same way, adding an infinite sequence before an infinite sequence of infinite sequences still gives us an infinite sequence of infinite sequences (i.e.
-$\omega +\omega^2 = \omega^2$).
-\end{remark}
-
-\begin{prop}[Cantor normal form]
-	Every ordinal number $\langle A \rangle$ can be written as a sum of powers of $\omega$.
-That is, $\langle A \rangle = \omega^{\alpha_1} c_1 + \omega^{\alpha_2} c_2 + ...
-+ \omega^{\alpha_n} c_n$ where $n \in \mathbb{N}$, $c_i \in \mathbb{N}^+$ for all $1 \leq i \leq n$ and $\alpha_i \in \ordinals$ such that $\alpha_i > \alpha_j$ for all $i > j$.
-\end{prop}
-
-\subsection{Cardinal numbers}
-
-\begin{remark}
-	The cardinality of a set (i.e.
-how many elements it contains) is defined in terms of injection/bijection: a set is smaller than another if it can be injected into it, and it is as big as another if it can be put into a one-to-one correspondence.
-\end{remark}
-
-\begin{defn}[Cardinal comparison]
-	Let $A$ \marginleft{Cardinality} and $B$ be two sets.
-If there exists a injective function $f : A \to B$ then we say $A$ \textbf{has lower cardinality than} $B$, noted $A \crdleq B$.
-If there exists a bijective function we say they have \textbf{the same cardinality}, noted $A \crdeq B$.
-\end{defn}
-
-\begin{prop}[Cardinal comparability of sets]
-	Given two sets $A$ and $B$, one must have lower cardinality than the other.
-That is, either $A \crdleq B$ or $B \crdleq A$.
-Additionally, there exists a bijective function between $A$ and $B$ if and only if there exists an injective function in both direction.
-That is, $A \crdeq B$ if and only if $A \crdleq B$ and $B \crdleq A$.
-\end{prop}
-
-\begin{prop}[Well-ordering theorem]
-	Every set can be well-ordered.
-\end{prop}
-
-\begin{remark}
-	The well-ordering theorem allows us to reuse ordinal numbers to quantify cardinality. As two different well-ordered sets may have the same cardinality, we'll take the lowest ordinal between them.
-\end{remark}
-
-\begin{defn}[Cardinal number]
-	A \textbf{cardinal}  \marginleft{Cardinals: $\cardinals$} is an ordinal $a \in \ordinals$ that has lower ordinality of any ordinal that has the same cardinality.
-That is, $a \ordleq b$ for all $b \in \ordinals$ such that $a \crdeq b$.
-The class of all cardinals is noted $\cardinals$ and, by definition $\cardinals \subset \ordinals$.
-\end{defn}
-
-\begin{prop}[Cardinality of a set]
-	For every \marginleft{$|A|$} set $A$ there is a unique cardinal that identifies the cardinality of $A$.
-That is, there exists a unique $|A| \in \cardinals$ such that $|A| \crdeq A$.
-We say $|A|$ is the \textbf{cardinality of $A$}.
-\end{prop}
-
-\begin{prop}[Cantor's theorem]
-	A set has strictly lower cardinality than its power set. That is, $|A| < |\mathcal{P}(A)|$. If $\alpha = |A|$ then we note $|\mathcal{P}(A)| = 2^\alpha$
-\end{prop}
-
-\begin{prop}[Aleph numbers]
-	The cardinality \marginleft{Aleph \\numbers: $\aleph_\alpha$} of infinite sets are well-ordered. They are noted $\aleph_\alpha$ where $\alpha \in \ordinals$.
-\end{prop}
-
-\begin{defn}[Aleph-nought]
-	The smallest infinite cardinal \marginleft{$\aleph_0$} $\aleph_0$ denotes countable infinity.
-\end{defn}
-
-\begin{defn}[Cardinality of continuum]
-	The cardinality \marginleft{Continuum: $2^{\aleph_0}$} of the continuum is the cardinality of the powerset of a countable set $2^{\aleph_0}$.
-\end{defn}
-
-\begin{remark}
-	Sets that have countable cardinality:
-	\begin{itemize}
-		\item natural $\mathbb{N}$, integer $\mathbb{Z}$ and rational $\mathbb{Q}$ numbers 
-		\item ordinals such as $\omega$, $\omega + 3$, $\omega2$, $\omega^\omega$
-	\end{itemize}
-	Sets that have cardinality of the continuum:
-	\begin{itemize}
-		\item powerset of natural, integers and rationals: $\mathcal{P}(\mathbb{N})$, $\mathcal{P}(\mathbb{Z})$, $\mathcal{P}(\mathbb{Q})$
-		\item real $\mathbb{R}$ and complex $\mathbb{C}$ numbers
-		\item real and complex euclidean spaces: $\mathbb{R}^n$, $\mathbb{C}^n$
-		\item continuous function over euclidean spaces: $C(\mathbb{R}^n, \mathbb{R}^m)$
-		\item standard topology and Borel algebra on $\mathbb{R}^n$ (i.e. the set of open sets and Borel set respectively)
-		\item countable sequences of numbers: $\mathbb{N}^\mathbb{N}$, $\mathbb{Z}^\mathbb{N}$, $\mathbb{Q}^\mathbb{N}$, $\mathbb{R}^\mathbb{N}$, $\mathbb{C}^\mathbb{N}$
-	\end{itemize}
-	Sets that have cardinality greater than the continuum:
-	\begin{itemize}
-		\item powerset of the real and complex numbers: $\mathcal{P}(\mathbb{R})$, $\mathcal{P}(\mathbb{C})$
-		\item set of all functions from $\mathbb{R}^n$ to $\mathbb{R}^m$
-	\end{itemize}
-
-	We should note that all physically meaningful sets have at most the cardinality of the continuum.
-This is not a coincidence.
-In science we are interested in object that can be experimentally identified.
-Every experimental test can acquire finite information in finite time and, assuming arbitrarily long time, we can run at most countably many tests.
-Countable sequences of numbers have the cardinality of the continuum, which is therefore the best we can distinguish.
-Even if nature allows for wider variety, it would be merged into equivalence classes of objects that are experimentally indistinguishable.
-Much of issues of higher cardinals can therefore be safely set aside by those working on the Assumptions of Physics.
-\end{remark}
-
-\begin{defn}[Aleph-one]
-	Let $\Omega$ \marginleft{$\aleph_1$} be the set of all countable ordinal numbers. Its cardinality $|\Omega| = \aleph_1$ is the second aleph number.
-\end{defn}
-
-\begin{remark}
-	 While it can be proven that the cardinality of the continuum $2^{\aleph_0}$, yet which one is undetermined.
-\end{remark}
-
-\textbf{Continuum hypothesis}
-	There is \marginleft{Continuum hypothesis} no infinite set with cardinality in between that of the continuum and countable infinity. That is: $2^{\aleph_0} = \aleph_1$.
-
-\begin{remark}
-	The continuum hypothesis is proven to be independent of the axioms of set theory.
-It can be taken to be true or to be false without inconsistency.
-As the axiom deals with sets of cardinality lower than the continuum, this aspect of set theory may be interesting for those working on the Assumptions of Physics.
-It could be, in principle, that the distinguishable configurations in space-time have a cardinality that is neither countable nor that of the continuum.
-\end{remark}
 

### 2021-12-09 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: minor changes

@@
 Therefore, according to the HS categorization, these correspond to two ontologically distinct states, making the preparation axis an ontological property of the system, regardless of the orientation.
-However, in terms of density matrixes both these situations would not just overlap: they would be represented by the same maximally mixed state.
-Hence there seems to be a mismatch between the type of epistemic knowledge one can express in quantum mechanics through mixed states and the type of epistemic knowledge described in the HS model.
+However, in terms of density matrixes both these situations would not just overlap: \emph{they would be represented by the same maximally mixed state.} Hence there seems to be a mismatch between the type of epistemic knowledge one can express in quantum mechanics through mixed states and the type of epistemic knowledge described in the HS model.
 This needs to be investigated further.

### 2021-12-09 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
BM-OrderTheory: first draft

\documentclass{article}

\usepackage[left=1.8in,right=1.8in,top=.6in,bottom=1in]{geometry}

\usepackage{assumptionsofphysics}
\usepackage{tikz}
\usetikzlibrary{positioning}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\frenchspacing

\newcommand{\marginleft}[1] {\reversemarginpar\marginpar{#1}}
\newcommand{\marginright}[1] {\normalmarginpar\marginpar{#1}}

\def\ordinals{\textbf{ORD}}
\def\cardinals{\textbf{CRD}}

\def\ordless{\prec}
\def\ordleq{\preceq}
\def\ordeq{\sim}
\def\ordgeq{\succeq}

\def\crdleq{\hookrightarrow}
\def\crdeq{\leftrightarrow}
\def\crdgeq{\hookleftarrow}

\title{Bare minimum: order theory}

\date{\vspace{-5ex}}
\begin{document}

\maketitle

\begin{abstract}
This note presents a condensed summary of order theory which can function as a crash course, refresher and/or reference. Bare minima are meant to give a rough overview, by no means complete, of the subject to the intellectually curious, particularly in the context of foundational questions in physics.
\end{abstract}

\section{Introduction}

The mathematical structures studied by order theory appear in different contexts, making the topic relevant to those interested in foundational questions. Core structures in fields such mathematical logic, topology and measure theory, can be characterized and understood in purely order theoretic terms; characterizing number types (i.e. integers, rational and reals) as different order types makes a better link to physics; physical concepts (e.g. part-whole inclusion) are also naturally captured with order theory.

Roughly speaking, orders are transitive relationships that are not symmetric. Order theory studies and catalogs these structures, together with order preserving transformations. For more details on order theory, see for example \cite{davey2002introduction, pinter2014book}.

\emph{This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.}

\section{TBD}

* fixpoint theorems
* Zorn's lemma

* Galois connection

? graded posets

\section{Orders}

\subsection{Basic definitions}

Mathematically, the following is the most convenient and common way to define order relationships.

\begin{defn}[Partial orders and poset]
	A \textbf{partial order} \marginleft{Partial order, poset: $(X, \leq)$} $\leq$ is a binary relationship over a set $X$ that is:
	\begin{description}
		\item[Reflexive] $a \leq a$ for all $a \in X$
		\item[Antisymmetric] if $a \leq b$ and $b \leq a$ then $a = b$ for all $a, b \in X$
		\item[Transitive] if $a \leq b$ and $b \leq c$ then $a \leq c$ for all $a, b, c \in X$.
	\end{description}
	A \textbf{partially ordered set}, or \textbf{poset}, is a tuple $(X, \leq)$ consisting on a set and a partial order.
\end{defn}

\begin{remark}
	To give a sense of the broad range of applicability: sets are ordered by set inclusion ($\{0, 1\} \subseteq \{0, 1, 2\}$); statements are ordered by implication/narrowness (\statement{Snoopy is a dog} $\narrower$ \statement{Snoopy is a mammal}); systems are ordered by is-a-part-of (`piston' $\subseteq$ `engine' $\subseteq$ `car').
\end{remark}

\begin{defn}[Linear order]
	A \textbf{linear order}, \marginleft{Linear order,\\ chain} or \textbf{total order}, is a partial order where every pair is ordered. That is, given $a, b \in X$ either $a \leq b$ or $b \leq a$. A \textbf{chain} typically refers to a subset of a partial order that is linearly ordered.
\end{defn}

The more intuitive notion of `before' and `after' would not include the equal. The two notions can be shows to be equivalent.

\begin{defn}[Strict order]
	A \textbf{strict partial order} \marginleft{Strict order} $<$ is a binary relationship over a set $X$ that is:
	\begin{description}
		\item[Irreflexive] $a \nless a$ for all $a \in X$
		\item[Asymmetric] if $a < b$ then $b \nless a$ for all $a, b \in X$
		\item[Transitive] if $a < b$ and $b < c$ then $a < c$ for all $a, b, c \in X$.
	\end{description}
\end{defn}

\begin{prop}
	A strict partial order can be converted to a non-strict partial order (or vice-versa) by adding (or removing) the identity relationship. That is, given $<$ we can define $\leq$ such that $a \leq b$ if $a < b$ or $a = b$. On the other hand, given $\leq$ we can define $<$ such that $a < b$ if $a \leq b$ and $a \neq b$.
\end{prop}

A more loose idea of order is a preorder, for which two different elements can be in the same order position. For example, the relationship has-no-fewer-elements for set is a preorder.

\begin{defn}[Preorder]
	A \textbf{preorder} \marginleft{Preorder} $\lesssim$ is a binary relationship over a set $X$ that is:
	\begin{description}
		\item[Reflexive] $a \lesssim a$ for all $a \in X$
		\item[Transitive] if $a \lesssim b$ and $b \lesssim c$ then $a \lesssim c$ for all $a, b, c \in X$.
	\end{description}
\end{defn}

\begin{prop}
	A preorder induces a partial order on the set of equivalence classes defined by the preorder. That is, define $\sim$ on $X$ such that $a \sim b$ if $a \lesssim b$ and $b \lesssim a$. This is an equivalence relationship. Now define $\lesssim$ on $X_{/\sim}$ such that $a_{/\sim} \lesssim b_{/\sim}$ if $a \lesssim b$. This is a preorder.
\end{prop}

\begin{defn}[Predecessors and successors]
	A \textbf{predecessor} \marginleft{Predecessor, successor} (or \textbf{successor}) of an element $a \in X$ is an element that comes strictly before (or strictly after) $a$. That is, it is an element $b \in X$ such that $b < a$ (or $a < b$). A predecessor (or successor) is  \textbf{immediate} if there are no elements in between. That is, $b < a$ (or $a < b$) and there is no $c \in X$ such that $b < c < a$  (or $a < c < b$).
\end{defn}

\marginright{TODO: Hesse diagram}

Simple orders \marginleft{Hasse diagram}  can be depicted using \textbf{Hasse diagrams}. Each element in the poset is represented by a vertex. A line connects upwards an element to its immediate successor, so that lowest elements are at the bottom.

\begin{defn}[Dense order]
	An order is \textbf{dense} \marginleft{Dense order} if for any two elements $a < b$ we can always find a third element $c \in X$ between the two $a < c < b$.
\end{defn}

\subsection{Morphisms}

\begin{defn}
	A \textbf{monotonic function} \marginleft{Monotonic function} (also called \textbf{monotone}, \textbf{increasing} or \textbf{order preserving}) is a map between two posets $(X, \leq_X)$ and $(Y, \leq_Y)$ that respects the order. That is, a map $f : X \to Y$ such that $a \leq_X b \implies f(a) \leq_Y f(b)$.
\end{defn}

\begin{defn}
	An \textbf{order embedding} \marginleft{Order embedding} is a monotonic function that respects the order in both directions. That is, a map $f : X \to Y$ such that $a \leq_X b \iff f(a) \leq_Y f(b)$.
\end{defn}

\begin{defn}
	An \textbf{order isomorphism} \marginleft{Order isomorphism} is an invertible embedding (i.e. bijective monotonic function whose inverse is monotonic). Two posets are \textbf{order isomorophic} if there exist an order isomoprhism between them.
\end{defn}

\begin{defn}
	Let $X$ be a poset. \marginleft{Up-sets and \\ down-sets} An \textbf{up-set}, or \textbf{increasing set}, or \textbf{order filter}, is a subset $A \subseteq X$ that contains all the successors of its elements. That is, if $a \in A$ then for all $b \in X$ such that $a \leq b$ we have $b \in A$. Dually, a \textbf{down-set}, or \textbf{decreasing set}, or \textbf{order ideal}, is a subset $A \subseteq X$ that contains all the predecessors of its elements. That is, if $a \in A$ then for all $b \in X$ such that $b \leq a$ we have $b \in A$.
\end{defn}

\subsection{Bounds}

\begin{defn}[Top and bottom]
	The \textbf{top} \marginleft{Top and bottom: $\top, \bot$} element of an ordered set $X$, noted $\top$ or $1$, is the greatest element, if it exists. That is, $\top \in X$ such that $a \leq \top$ for all $a \in X$. Dually, the \textbf{bottom} element, noted $\bot$ or $0$, is the least element, if it exists. That is, $\bot \in X$ such that $\bot \leq a$ for all $a \in X$. An order that admits both a top and a bottom is said \textbf{bounded}. Conversely, an order that admits neither is said \textbf{unbounded}.
\end{defn}

\begin{defn}[Upper and lower bounds]
	Let $P$ be an \marginleft{Upper/Lower bounds} ordered set and let $A \subseteq P$. An \textbf{upper bound} (or \textbf{lower bound}) is an element $p \in P$ such that $a \leq p$ (or $p \leq a$) for all $a \in A$.
\end{defn}

\begin{defn}[Supremum and infimum]
	The \textbf{supremum}, \marginleft{Supremum and infimum} or \textbf{least upper bound} of $A$ is, if it exists, the least element within all upper bounds of $A$. Conversely, the \textbf{infimum},  or \textbf{greatest lower bound} of $A$ is, if it exists, the greatest element within all lower bounds of $A$.
\end{defn}

\begin{defn}[Join and meet]
	The \textbf{join} \marginleft{Join and meet: $\vee, \wedge$} of two elements $a, b \in X$ or of a set $A$, noted $a \vee b$ and $\bigvee A$, is their supremum if it exists. The \textbf{meet} of two elements $a, b \in X$, noted $a \wedge b$, is their infimum, if it exists.
\end{defn}

\begin{defn}[Complement]
	Let $X$ be \marginleft{Complement: $\neg$} a bounded ordered set. Two elements $a, b \in X$ are \textbf{complements} of each other if $a \vee b = \top$ and $a \wedge b = \bot$. If an element $a$ has a unique complement, it is noted as $\neg a$.
\end{defn}

\begin{defn}[Lattice - order theoretic]
	A \textbf{lattice} \marginleft{Lattice} is a partially ordered sets in which every pair of elements has both a supremum/join and an infimum/meet.
\end{defn}

\begin{prop}[Connecting lemma]
	If $X$ is a lattice then $a \leq b$, $a \vee b = b$ and $a \wedge b = a$ are equivalent.
\end{prop}

\begin{prop}
	Every linearly ordered set is a lattice.
\end{prop}

\begin{defn}[Lattice properties]
	We define\marginleft{Complete, distributive, complemented} the following additional properties on lattices:
	\begin{description}
		\item[Completeness.] A lattice is \textbf{complete} if all subsets have joins and meet, not just finite ones.
		\item[Distributivity.] A lattice is \textbf{distributive} if it satisfies
		\begin{align*}
			a \vee (b \wedge c) = (a \vee b) \wedge (a \vee c) \\
			a \wedge (b \vee c) = (a \wedge b) \vee (a \wedge c).
		\end{align*}
		A complete lattice is \textbf{infinitely distributive} if the same holds over infinite operations.
		\item[Complemented.] A bounded lattice is \textbf{complemented} if every element has a unique complement.
	\end{description}
\end{defn}

\begin{prop}
	Every lattice \marginleft{Completion} admits a \textbf{completion}, the smallest complete lattice that contains it. The completion is unique up to an order isomorphism.
\end{prop}

\begin{remark}
	The completion is constructed by finding pairs $(A, B)$ of subsets $A, B \subseteq X$ such that $A$ is the set of lower bounds for $B$ and $B$ is the set of lower bounds of $A$. For the rationals $\mathbb{Q}$, these pairs correspond to Dedekind cuts, and the completion of $\mathbb{Q}$ as a lattice is the lattice of the reals $\mathbb{R}$.
\end{remark}

\begin{prop}[Zorn's lemma]
	Every non-empty partially ordered set for which every linearly ordered subset has an upper bound contains at least one maximal element.
\end{prop}

\begin{defn}[Maximal subsets]
	A linearly \marginleft{Maximal subset} ordered subset is \textbf{maximal} if it is not a scrict subset of another linearly ordered subset.
\end{defn}

\begin{prop}[Kuratowski's Lemma]
	Every linearly ordered subset is the subset of some maximal linearly ordered subset.
\end{prop}

\section{Algebras and orders}

\begin{defn}[Lattice - algebraic]
	An algebraic \marginleft{Lattice (algebraic)} structure $(X, \vee, \wedge)$ consisting of a set $X$ and two binary operations $\vee$ and $\wedge$ is an \textbf{(algebraic) lattice} if it satisfies the following properties:
	\begin{description}
		\item[Associativity.] $(a \vee b) \vee c = a \vee (b \vee c)$
		and $(a \wedge b) \wedge c = a \wedge (b \wedge c)$
		\item[Commutativity.] $a \vee b = b \vee a$
		and $a \wedge b = b \wedge a$
		\item[Idempotency.] $a \vee a = a$
		and $a \wedge a = a$
		\item[Absorption.] $a \vee (a \wedge b) = a$
		and $a \wedge (a \vee b) = a$
	\end{description}
\end{defn}

\begin{prop}
	Every order lattice is an algebraic lattice and every algebraic lattice is an order lattice where $a \leq b$ if $a \vee b = b$.
\end{prop}

\begin{defn}[Boolean algebra]
	An algebraic \marginleft{Boolean algebra} structure $(X, \vee, \wedge, \neg, \bot, \top)$ is a \textbf{Boolean algebra} if it satisfies the following properties:
	\begin{description}
		\item[Lattice.] $(X, \vee, \wedge)$ is a lattice
		\item[Distributivity.] $a \vee (b \wedge c) = (a \vee b) \wedge (a \vee b)$
and $a \wedge (b \vee c) = (a \wedge b) \vee (a \wedge b)$
		\item[Identity.] $a \vee \bot = a$
		and $a \wedge \top = a$
		\item[Complementation.] $a \vee \neg a = \top$
		and $a \wedge \neg a = \bot$
	\end{description}
\end{defn}

\begin{prop}
	Every complemented distributive lattice is a Boolean algebra and every Boolean algebra is a complemented distributive lattice.
\end{prop}

\begin{defn}
	Let $X$ be \marginleft{Pseudo-complement} a bounded lattice. The \textbf{relative pseudo-complement} of $a$ with respect to $b$ is the greatest element $x$ such that $a \wedge x \leq b$. The \textbf{pseudo-complement} of $a$ is the relative pseudo-complement with respect to $\bot$.
\end{defn}

\begin{defn}
	A \textbf{Heyting algebra} \marginleft{Heyting algebra} is a bounded lattice for which the relative pseudo-complement always exists.
\end{defn}

\begin{prop}
	All Heyting algebras are distributive lattices. Every Boolean algebra is a Heyting algebra where the pseudo-complement coincide with the complement. If a Heyting algebra is complemented then it is a Boolean algebra.
\end{prop}

\begin{remark}
	Boolean algebras are used to capture the two-valued (i.e. true, false) classical logic. Heyting algebras are used to capture the three-valued (i.e. true, false, undetermined) intuitionistic/constructivist logic. The order is the logical implication. The complement is the negation, which always exists in classical logic but not in intuitionistic (i.e. not true is not necessarily false).
\end{remark}

\section{Representation as sets}

\begin{prop}[Stone's representation theorem]
	Every distributive lattice is order isomorphic to a lattice of subsets ordered by inclusion.
\end{prop}

\begin{prop}[Boolean algebras]
	Every Boolean algebra can be represented by a collection of sets ordered by inclusion with the following correspondence:
	\begin{description}
		\item[Join] represented by the union
		\item[Meet] represented by the intersection
		\item[Pseudo-complement] represented by the set complement.
	\end{description}
	Conversely, every collection of sets closed under union, intersection and complement forms a Boolean algebra.
\end{prop}

\begin{prop}
	Every distributive lattice (or a Heyting algebra in particular) is a subalgebra a Boolean algebra since it is order isomorphic to a sublattice of a power set.
\end{prop}

\begin{prop}[Topology]
	Every topology, as a collection of sets ordered by inclusion, is a complete Heyting algebra with the following correspondence:
	\begin{description}
		\item[Join] represented by arbitrary union
		\item[Meet] represented by the interior of arbitrary intersection
		\item[Pseudo-complement] represented by the set exterior.
	\end{description}
\end{prop}

\begin{prop}[Lattice of subgroups]
	Given a group, the set of its subgroups, as a collection of sets ordered by inclusion, is a lattice following correspondence:
	\begin{description}
		\item[Join] represented by the subgroup generated by the union
		\item[Meet] represented by the intersection.
	\end{description}
\end{prop}

\begin{prop}[Lattice of subspaces]
	Given a vector space, the set of its subspaces, as a collection of sets ordered by inclusion, is a complemented lattice following correspondence:
	\begin{description}
		\item[Join] represented by the subspace spanned by the union
		\item[Meet] represented by the intersection
		\item[Complement] represented by the orthogonal subspace.
	\end{description}
\end{prop}

\begin{remark}
	Note how the order theoretic operations (i.e. join, meet, complement, ...), except for Boolean algebras, do not always map to the set theoretic operations (i.e. intersection, union, complement, ...). This creates confusion when one tries to attach intuition from one to the other.

	In Heyting algebras the pseudo-complement is often called ``negation'' and noted as $\neg a$ because, in some formulas, plays the same formal role. However, it is not a negation in a strict sense since the join of an element with its psuedo-complement is not the top. Logically, the negation of ``proposition $a$ is proven to be true'' is ``proposition $a$ is not proven to be true'' while the pseudo-complement is ``proposition $a$ is proven to be false''.

	Similarly, quantum logic takes the lattice of subspaces as the lattice of propositions. The lattice fails to be distributive on the join operation, and this is taken to mean that the disjunction is not distributive. However, the join fails to map to the union. Logically, the disjunction of \statement{the spin of system was prepared in the z+ direction} and \statement{the spin of system was prepared in the z- direction} is \statement{the spin of system was prepared in either the z+ or z- direction} while the join in terms of subspaces is \statement{the spin of system was prepared in any direction}.

	Understanding these differences is critical to give the correct physical interpretation of the corresponding mathematical definitions and results.

\end{remark}

\section{Numbers as order types}

\begin{prop}
	Every linearly ordered set $X$ in which each element has one predecessor and one successor is order isomorphic to the integers $\mathbb{Z}$.
\end{prop}

\begin{prop}
	 Every unbounded countable dense linearly ordered set is order isomorphic to the rationals $\mathbb{Q}$.
\end{prop}

\begin{prop}
	Every complete unbounded dense linearly ordered set which admits a countable dense subset is order isomorphic to the reals $\mathbb{R}$.
\end{prop}

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

### 2021-12-14 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: abstract

@@
 \begin{abstract}
-We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics.
-The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program.
-This allows us to identify physical starting points (as opposed to mathematical ones), compare their deductive power, and find those that can better serve as a foundation for our theories.
+The ontological model framework introduced by Harrigan and Spekkens has been widely used to categorize quantum interpretation and as a basis for influential results such as the PBR theorem.
+In this paper we show that the model has a fundamental flaw: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics.
+Namely, the map between the epistemic states of the model and quantum density matrices is not even an order homomorphism in terms of the information entropy, which means the epistemic content of mixed states is not mapped in a meaningful way.
+Moreover, by recasting the definitions of the model directly in terms of density matrices, we find that only $\psi$-epistemic interpretations are allowed by quantum mechanics, a result that is the opposite of the PBR theorem.
+We show that the cause of this mismatch is the effect of non-contextuality over mixed preparations that does not allow for a single decomposition of mixed states over pure states.
 \end{abstract}
@@
 
-Let us first characterize the extent of the discepancy.
+Let us first characterize the extent of the discrepancy.
 In HS, the space of epistemic states $E_{HS}$ is the set of probability distributions over pure states, which is an infinite dimensional space.
-This is a direct consequence of defining the epistemic states as a distribution $P(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability to each ontological state $\lambda$.\footnote{TODO: HS seems to be very vague about probability and probability density.} In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
+This is a direct consequence of defining the epistemic states as a distribution $p(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability $p$ to each ontological state $\lambda$.\footnote{TODO: HS seems to be very vague about probability and probability density.} In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
 Looking at the difference in dimension, the discrepancy is not minor.
@@
 Also note that all epistemic states in both spaces can be reached by mixture of pure states.
-The difference is that, in the quantum case, we have an ``aliasing'': different mixtures will give the same mixed states.
-Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{Q}$ that allows us to go from HS epistemic states to density matrixes.
+The difference is that, in the quantum (XXX) case, we have an ``aliasing'': different mixtures will give the same mixed states.
+Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{Q}$ that allows us to go from HS epistemic states to density matrices.
 This map is surjective but not injective.

### 2021-12-31 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: section 2 and figures added

--- 2021-12-01 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
@@
-\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
-%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
-%\documentclass[10pt,twocolumn,letterpaper]{article}
+%\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
+\documentclass[twocolumn,prl,floatfix,superscriptaddress]{revtex4-2}
 
-\RequirePackage{amsthm}
-\RequirePackage{amsmath}
-\RequirePackage{amssymb}
-\RequirePackage{enumitem}
+\usepackage{amsthm, amsmath, amssymb, enumitem, graphicx, hyperref}
+\usepackage{graphicx, subfigure}
 \setlist{noitemsep}
-\usepackage{graphicx}
-\usepackage{hyperref}
 \hypersetup{
@@
 \urlstyle{same}
+
+%\usepackage{graphicx, subfigure, amsfonts,amssymb,amsmath, array, gensymb,fontenc,color,lipsum}
+%\documentclass[aps,twocolumn,showpacs,preprintnumbers]{revtex4}
+%\usepackage{graphicx}  % Include figure files
+%\usepackage{subfigure}
+%\usepackage{multirow}
+%\linespread{1.0}
+%\usepackage{fancyhdr}
+%\usepackage{longtable}
+%\usepackage{parskip}
+%\usepackage[T1]{fontenc}
+%\usepackage{dcolumn}
+%\usepackage{bm}        
+%\usepackage{amsfonts}  
+%\usepackage{amsmath}   
+%\usepackage{amssymb}
+%\usepackage{hyperref}
+%\hypersetup{colorlinks= true, allcolors=blue}
+%\setcitestyle{aysep={}}
+%\setlength{\parindent}{11pt}
 \frenchspacing
-
-\newcommand\partitle[1]{\textsc{#1}.}
 
@@
 \title{On the reality of the quantum state once again}
-\author{Gabriele Carcassi, Andrea Oldofredi, Christine A. Aidala}
+\author{Gabriele Carcass}
 \affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\author{Andrea Oldofredi}
+\affiliation{Department of Philosophy, University of Lisbon, Portugal}
+\author{Christine Aidala}
+\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\vspace{2mm}
 
@@
 \begin{abstract}
-We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics.
-The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program.
-This allows us to identify physical starting points (as opposed to mathematical ones), compare their deductive power, and find those that can better serve as a foundation for our theories.
+The ontological model framework introduced by Harrigan and Spekkens has been widely used to categorize quantum interpretations and as a basis for influential results such as the PBR theorem.
+In this paper we show that the model has a fundamental flaw: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics.
+Namely, the map between the epistemic states of the model and quantum density matrices is not even an order homomorphism in terms of the information entropy, consequently the epistemic content of mixed states is not mapped in a meaningful way.
+Moreover, by recasting the definitions of the model in terms of density matrices, we find that only $\psi$-epistemic interpretations are allowed by quantum mechanics, a result that is the opposite of the PBR theorem.
+We show that the cause of this mismatch is the effect of non-contextuality over mixed preparations that does not allow for a single decomposition of mixed states over pure states.
 \end{abstract}
@@
 \section{Introduction}
-%(mezza pagina)
+
 %- Status of field and impact of HS
+%- Mention the PBR Theorem explicitly
 %- Claim that HS does not work for quantum epistemic cases, therefore invalidating all results
 
-\section{Summary of the model}
-%(una pagina)
-%- Ontic/supplemented/epistemic
+\section{Summary of the Harrigan \& Spekkens Model}
+
+\subsection{Ontological Models}
+
+Harrigan and Spekkens (HS) provide a classification of ontological quantum models that essentially depends on the nature of the quantum state \cite{Harrigan:2010}.
+Such ontological models are defined by employing an operational setting where the primitive of descriptions consist in preparation protocols and measurements performed on quantum systems.
+More precisely, HS claim that preparations procedures are ``assumed to prepare a system with certain properties and a measurement procedure is assumed to reveal something about those properties'' (\cite{Harrigan:2010}, p.\ 128).
+In their approach the complete specification of the attributes of a given physical object is provided by $\lambda$, the ontic state of the system under scrutiny.
+Such operational models provide the probabilities $p(k|M, P)$ to obtain outcomes $k$ for some measurement $M$ performed on the prepared systems, given the set of preparation instructions $P$.
+Indeed, from a formal perspective every protocol $P$ is associated with a density operator $\rho$ on the relevant Hilbert space, and every measurement $M$ is associated with a POVM $\{ E_k\}$.
+As usual, each measurement result $k$ can be obtained with probability $p(k|M, P)=\textrm{Tr}(\rho E_k)$, the generalized Born rule.
+To this regard, HS underline that although an agent perfectly knows the preparation procedures prior the performance of a certain measurement, they may have incomplete knowledge of $\lambda$.
+This is means that an observer which has incomplete information about $\lambda$ assigns ``non-sharp'', i.e.\ overlapping probability distributions $p(\lambda| P)$ over $\Lambda$, the space of ontic states.
+
+We can eventually provide the formal definition of ontological models as follows (\cite{Harrigan:2010}, p.\ 128): ``An ontological model of operational quantum theory posits an ontic state space $\Lambda$ and prescribes a probability distribution over $\Lambda$ for every preparation procedure $P$, denoted $p(\lambda| P)$, and a probability distribution over the different outcomes $k$ of a measurement $M$ for every ontic state $\lambda \in \Lambda$, denoted $p(k|\lambda, M)$.
+Finally, for all $P$ and $M$, it must satisfy
+\begin{align*}
+\int d\lambda p(k|\lambda, M) p(\lambda| P)= \textrm{tr}(\rho E_k),
+\end{align*}
+\noindent where $rho$ is the the density operator associated with $P$ and $E_k$ is the POVM element associated with outcome $k$ of $M$''.
+
+\subsection{$\psi-$ontic and $\psi-$epistemic models}
+
+Let us now introduce the first categorization of ontological models.
+HS divide them in $\psi-$\emph{complete} and $\psi-$\emph{incomplete}: in the former case, the quantum state encodes every information about the physical system at hand, as in the case of standard Quantum Mechanics (QM).
+Alternatively stated, in $\psi$-complete models there is a one-to-one relation between pure quantum states and the ontic state of physical items they represent.
+Formally, this is translated into the following definition (\cite{Harrigan:2010}, p.\ 131): ``An ontological model is $\psi-$complete if the ontic state space $\Lambda$ is isomorphic to the projective Hilbert space $\mathcal{PH}$ (the space of rays of Hilbert space) and if every preparation procedure $P_{\psi}$ associated in quantum theory with a given ray $\psi$ is associated in the ontological model with a Dirac delta function centered at the ontic
+state $\lambda_{\psi}$ that is isomorphic to $\psi$, $p(\lambda|P)=\delta(\lambda-\lambda_{\psi})$''.
+In the latter case, $\psi$ alone is not sufficient to provide a complete description of reality.
+Models in which $\psi$ is supplemented with additional/hidden variables are defined by HS as $\psi-$supplemented (well-known examples are Bohm pilot-wave theory \cite{Bohm:1952aa} and Nelson mechanics \cite{Nelson:1966aa}).
+
+In the second place, the authors propose another distinction concerning the nature of quantum states, classifying $\psi$ either as \emph{ontic} or \emph{epistemic}.
+In a nutshell, a model is said to be $\psi-$ontic if ``for any pair of preparation procedures, $P_{\psi}$ and $P_{\phi}$, associated with distinct quantum states $\psi$ and $\phi$, we have $p(\lambda | P_{\psi})p(\lambda|P_{\phi})=0$ for all $\lambda$'' (\cite{Harrigan:2010}, pp.\ 131-132).
+Consequently, (i) in $\psi-$ontic models the ontic state $\lambda$ can be consistently described only by a unique pure quantum state $\psi$, and (ii) observers' epistemic states associated with different quantum states are non-overlapping in $\Lambda$, i.e.\ they correspond to disjoint regions of the space of ontic states.
+Alternatively, in $\psi-$ontic models different pure quantum states correspond to disjoint probability distributions $\mu$ over $\Lambda$ (cf.\ Figure 1a).
+
+If a model in not $\psi-$ontic, then it is $\psi-$epistemic.
+A model is defined $\psi-$epistemic if $\lambda$ can be described by more than one quantum state.
+Therefore, in such epistemic models there are quantum states that correspond to overlapping probability distributions over $\Lambda$ (cf.\ Figure 1b).
+This fact implies that agents' epistemic states overlap, i.e.\ there exist preparation procedures $P_{\psi}, P_{\phi}$ and a $\lambda\in\Lambda$ such that $p(\lambda | P_{\psi})p(\lambda|P_{\phi})\neq0$.
+In turn, this entails that the ontic state $\lambda$ can be consistently represented by both quantum states $\psi$ and $\phi$: ``[i]n a $\psi-$epistemic model, multiple distinct quantum states are consistent with the same state of reality---the ontic state $\lambda$ does not encode $\psi$'' (\emph{ibid.}, p.\ 132).
+This explains why the authors claim that in $\psi-$epistemic models quantum states refer to observers' incomplete knowledge of reality, and not to reality itself.
+
+\begin{figure}
+\includegraphics[scale=.7]{ontic}
+\caption{\footnotesize{Harrigan and Spekkens distinction between $\psi-$ontic (a) and $\psi-$epistemic (b) ontological models.}}
+\end{figure}
+
+Now we can combine both distinctions in order to classify different ontological models.
+As we have seen, in $\psi-$complete models there is a one-to-one relation between $\lambda$ and its description provided by $\psi$.
+Consequently, if one knows the quantum state $\psi$ then one has a complete knowledge of the ontic state $\lambda$ of the system under consideration.
+In this models a variation of $\psi$ implies a variation of the ontic state, therefore $\psi-$complete models are $\psi-$ontic.
+If ontological models are $\psi-$incomplete, then they may be either $\psi-$supplemented or $\psi-$epistemic.
+In the former case the description of a physical system provided by $\psi$ is incomplete and is supplemented with additional (or hidden) variables whose value is generally unknown.
+Notably, also the class of $\psi-$supplemented models is $\psi-$ontic.
+Finally, in $\psi-$epistemic models $\psi$ represents an agent's incomplete knowledge of reality (of $\lambda$), and not reality itself; in this case a variation of $\psi$ does not necessarily entail a variation of $\lambda$.
+
+From these distinctions some conclusions can be inferred.
+In the first place, models which are simultaneously $\psi-$complete \emph{and} $\psi-$epistemic cannot exist.
+Therefore, if a model is $\psi-$complete, it must be $\psi-$ontic (cf.\ Lemma 6, \cite{Harrigan:2010}, p.\ 133).
+Alternatively, if a model is $\psi-$incomplete, then it can either be $\psi-$ontic, as in the case of $\psi-$supplemented models, or $\psi-$epistemic.
+If a model is $\psi-$epistemic, then it cannot be $\psi-$ontic, since it does not describe any underlying physical reality, but only the agents' knowledge of it.
 
@@
 
-Let us first characterize the extent of the discepancy.
+Let us first characterize the extent of the discrepancy.
 In HS, the space of epistemic states $E_{HS}$ is the set of probability distributions over pure states, which is an infinite dimensional space.
@@
 The key question, then, is the following: is the map $\iota$ sufficiently well behaved that we can understand the epistemic content of $E_{Q}$ based on the epistemic content of $E_{HS}$?
-Or is it so hopelessy broken that there is no connection between the two?
+Or is it so hopelessly broken that there is no connection between the two?
 
@@
 This is inconsistent with the idea that $E_{Q}$ represents states where knowledge is lost: that would pick the maximum possible entropy.
+
+\begin{figure}
+\includegraphics[scale=.4]{fig2}
+\caption{\footnotesize{}}
+\end{figure}
 
@@
 
-\section*{Supplementary material}
+%\section*{Supplementary material}
+%
+%TODO: do we need to bump anything here?
 
-TODO: do we need to bump anything here?
-
-\bibliography{bibliography}
+\bibliographystyle{apalike}
+\bibliography{PhDthesis}
 

### 2022-01-06 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: review with Christine

--- 2021-12-14 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
@@
 \title{On the reality of the quantum state once again}
-\author{Gabriele Carcassi, Andrea Oldofredi, Christine A. Aidala}
+\author{Gabriele Carcassi}
 \affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\author{Andrea Oldofredi}
+\affiliation{Department of Philosophy, University of Lisbon, Portugal}
+\author{Christine Aidala}
+\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\vspace{2mm}
 
@@
 \begin{abstract}
-The ontological model framework introduced by Harrigan and Spekkens has been widely used to categorize quantum interpretation and as a basis for influential results such as the PBR theorem.
+The ontological model framework introduced by Harrigan and Spekkens has been widely used to categorize quantum interpretations and as a basis for influential results such as the PBR theorem.
 In this paper we show that the model has a fundamental flaw: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics.
-Namely, the map between the epistemic states of the model and quantum density matrices is not even an order homomorphism in terms of the information entropy, which means the epistemic content of mixed states is not mapped in a meaningful way.
-Moreover, by recasting the definitions of the model directly in terms of density matrices, we find that only $\psi$-epistemic interpretations are allowed by quantum mechanics, a result that is the opposite of the PBR theorem.
+Namely, the map between the epistemic states of the model and quantum density matrices is not even an order homomorphism in terms of the information entropy, consequently the epistemic content of mixed states is not mapped in a meaningful way.
+Moreover, by recasting the definitions of the model in terms of density matrices, we find that only $\psi$-epistemic interpretations are allowed by quantum mechanics, a result that is the opposite of the PBR theorem.
 We show that the cause of this mismatch is the effect of non-contextuality over mixed preparations that does not allow for a single decomposition of mixed states over pure states.
@@
 \section{Introduction}
-%(mezza pagina)
+
 %- Status of field and impact of HS
+%- Mention the PBR Theorem explicitly
 %- Claim that HS does not work for quantum epistemic cases, therefore invalidating all results
 
-\section{Summary of the model}
-%(una pagina)
-%- Ontic/supplemented/epistemic
+TODO intro with status of the field
 
-\section{Quantum epistemic states}
-Now that we have presented the model, we want to study the space of epistemic states.
+In this paper we will show two things:
+\begin{enumerate}
+	\item the information entropy as calculated on epistemic states does not match the one provided by quantum mechanics, meaning that the two epistemic structures are incompatible;
+	\item the density matrices associated with pure states overlap for non-orthogonal states (i.e. $\rho_\psi \rho_\phi \neq 0$), meaning $\psi$-ontic models are ruled out.
+\end{enumerate}
+Our conclusion is that the ontological model is fundamentally flawed.
+The problem lies in assuming that the epistemic state identifies a single probability distribution over the whole set of ontic states.
+We show that this, in essence, is an assumption of non-contextuality.
+
+\section{Summary of the Harrigan \& Spekkens Model}
+
+We briefly review the main features of the classification provided by Harrigan and Spekkens (HS)\cite{Harrigan:2010}.
+We start with the usual operational setting for quantum mechanics.
+We have a preparation protocol $P$, which will be associated with a density operator $\rho$ on the relevant Hilbert space, and a measurement protocol $M$, represented by a POVM $\{ E_k\}$ where each $k$ represents a possible measurement outcome.
+The probability of obtaining a particular $k$ given a particular preparation $P$ and measurement $M$ is given by the generalized Born rule
+\begin{equation}
+	p(k|M, P)=\textrm{tr}(\rho E_k).
+\end{equation}
+
+An ontological model, as defined by HS, additionally assumes that there exists a set of states $\Lambda$, called \textbf{ontic states}, that provide the complete specification of a given physical object.
+A preparation $P$ will prepare a particular ontic state $\lambda$ according to the probability distribution $p(\lambda | P)$, which is referred to as \textbf{epistemic state}.
+The measurement outcome will depend only on the ontic state with probability $p(k|\lambda, M)$.(\cite{Harrigan:2010}, p.\ 128).
+This leads to the following expression:
+\begin{equation}
+	\int_\Lambda d\lambda p(k|\lambda, M) p(\lambda| P)= \textrm{tr}(\rho E_k).
+\end{equation}
+It is also assumed that a mixture of pure states $\{ \psi_i \}$ with probabilities $\{ w_i \}$ will be represented by
+\begin{equation}\label{epistemic_mixing}
+	\sum_i  w_i p(\lambda| P_{\psi_i}).
+\end{equation}
+As the epistemic states for mixed preparations are simply linear combinations of epistemic states for pure preparations, we can concentrate on the latter.
+
+To classify an ontological model, we look at the relationship between quantum and ontic states.
+There are two broad categories.
+In a \textbf{$\psi$-ontic} model the wave function is a physical property of the ontic state, in the sense that given an ontic state $\lambda$ there is only one pure state preparation $P_\psi$ that could have prepared $\lambda$.
+As we can see in figure \ref{overlap}, this happens if the probability distributions do not overlap, i.e.
+if we have
+\begin{equation}\label{ontic_condition}
+	p(\lambda | P_{\psi})p(\lambda|P_{\phi})=0
+\end{equation}
+for all pairs of states $\psi$ and $\phi$.
+If a model is not $\psi$-ontic, it is called \textbf{$\psi$-epistemic} and the wave function is taken to represent knowledge about the state preparation.
+A $\psi$-ontic model is said \textbf{$\psi$-complete} if the quantum states and the ontic states coincide.
+More precisely if
+\begin{equation}\label{complete_condition}
+	p(\lambda|P_\psi)=\delta(\lambda-\lambda_{\psi}).
+\end{equation}
+All other models are \textbf{$\psi$-incomplete}.
+
+To make this classification more concrete, let us give an example from classical mechanics.
+Consider the case where we prepare a particle according to a specific value of energy.
+The energy partitions phase space into mutually exclusive regions, and therefore we can understand the energy of the preparation as a property of the particle itself.
+According to HS, this would be an \emph{ontic property}.
+Consider now the case where we prepare a particle according to a specific temperature.
+When we take the particle from the oven, we are sampling from a Boltzmann distribution over different energies.
+Unlike energy, temperature does not partition phase space because the same particle state could have been prepared by ovens at different temperature.
+Temperature is a property of the preparation and therefore an \emph{epistemic property}.
+
+While the original aim of this classification was to clarify Einstein's view of quantum theory, it has been used in the literature to classify different interpretations (e.g.
+hidden variable theories such as Bohm pilot-wave theory \cite{Bohm:1952aa} and Nelson mechanics \cite{Nelson:1966aa} are both considered $\psi$-ontic and $\psi$-incomplete) and also to argue what types of interpretations are possible.
+Most notably, the PBR theorem rules out any $\psi$-epistemic model (TODO beef up this paragraph with citations).
+It is therefore worthwhile to look at the ontological model itself with more scrutiny.
+
+\begin{figure}
+\includegraphics[scale=.7]{ontic}
+\caption{\footnotesize{Harrigan and Spekkens distinction between $\psi-$ontic (a) and $\psi-$epistemic (b) ontological models.
+TODO: change the notation to $p(\lambda | P_\psi)$; use product of probability equal to zero}}\label{overlap}
+\end{figure}
+
+\section{Epistemic structure}
+
+We look more closely at the space of epistemic states given by $p(\lambda | P)$ and equation \ref{epistemic_mixing}.
 To make the discussion more concrete, we will assume we have a $\psi$-complete ontological model of a single qubit representing the direction of spin.
 
-Suppose we have two preparation devices, one that prepares z+ and another that prepares z-, and we toss a fair coin to choose which one to use.
-This results into an epistemic state with equal distribution on z+ and z-.
-Now suppose we have two additional preparation devices, this time for x+ and x-, and we again toss a fair coin.
+Suppose we have a protocol that prepares either z+ or z- with equal chance.
+This results in an epistemic state with equal distribution on z+ and z-.
+Now suppose we have another protocol that prepares x+ and x- again with equal chance.
 This time we have an epistemic state with equal distribution on x+ and x-.
@@
 Therefore, according to the HS categorization, these correspond to two ontologically distinct states, making the preparation axis an ontological property of the system, regardless of the orientation.
-However, in terms of density matrixes both these situations would not just overlap: \emph{they would be represented by the same maximally mixed state.} Hence there seems to be a mismatch between the type of epistemic knowledge one can express in quantum mechanics through mixed states and the type of epistemic knowledge described in the HS model.
+However, the respective density matrices would not just overlap: they would be the same maximally mixed state.
+Hence there seems to be a mismatch between the type of epistemic knowledge one can express in quantum mechanics through mixed states and the type of epistemic knowledge described in the HS model.
 This needs to be investigated further.
@@
 In HS, the space of epistemic states $E_{HS}$ is the set of probability distributions over pure states, which is an infinite dimensional space.
-This is a direct consequence of defining the epistemic states as a distribution $p(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability $p$ to each ontological state $\lambda$.\footnote{TODO: HS seems to be very vague about probability and probability density.} In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
+This is a direct consequence of defining the epistemic states as a distribution $p(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability to each ontological state $\lambda$.\footnote{We will follow the convention of not distinguishing between probability and probability density.} In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
 Looking at the difference in dimension, the discrepancy is not minor.
@@
 Also note that all epistemic states in both spaces can be reached by mixture of pure states.
-The difference is that, in the quantum (XXX) case, we have an ``aliasing'': different mixtures will give the same mixed states.
-Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{Q}$ that allows us to go from HS epistemic states to density matrices.
+The difference is that, in the quantum case, we have an ``aliasing'': different mixtures will give the same mixed states.
+Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{Q}$ that allows us to go from HS-epistemic states to density matrices, the Q-epistemic states.
 This map is surjective but not injective.
@@
 The key question, then, is the following: is the map $\iota$ sufficiently well behaved that we can understand the epistemic content of $E_{Q}$ based on the epistemic content of $E_{HS}$?
-Or is it so hopelessy broken that there is no connection between the two?
+Or is it so ill behaved that there is no connection between the two?
 
-Ideally, one would like to say is that $E_{HS}$ represents all possible epistemic states that include the knowledge about preparation, while $E_{Q}$ represents only those distinguishable through measurement.
+Ideally, one would like to say that $E_{HS}$ represents all possible epistemic states that include the knowledge about preparation, while $E_{Q}$ represents only those distinguishable through measurement.
 For example, one may want to say that the axis of preparation is, in principle, an ontologically distinct property, but this is not experimentally distinguishable.
-Therefore distinct epistemic states for each direction are projected by $\iota$ onto the same epistemic state of maximum uncertainty.
+Therefore distinct HS-epistemic states for each direction are projected by $\iota$ onto the same Q-epistemic state of maximum uncertainty.
 This sounds very plausible and appealing, but it turns out this is not at all the case.
-To show this we will concentrate on a single aspect about the map: how information entropy transforms under the map.
-While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is sufficient to claim that the epistemic structures of $E_{HS}$ and $E_{Q}$ are different.
+To show this we will concentrate on a single aspect of the map: how information entropy transforms under the map.
+While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is sufficient to show that the epistemic structures of $E_{HS}$ and $E_{Q}$ are different.
 
-Since every pure state in quantum mechanics has zero entropy, we will assume that this is the case for all pure states in $E_{HS}$.
-Let us consider three uniform distributions over a sphere with support given by:
+Guided by figure \ref{fig_map}, let us study the map $\iota$ by going through a few examples.
+Since every pure state in quantum mechanics has zero entropy, we will require the same for all pure states in $E_{HS}$.
+Let us consider the HS-epistemic states specified by three uniform distributions over the Bloch sphere with support given by:
 \begin{description}
@@
 \end{description}
-Since we want the entropy for a single microstate to be zero, the entropy for A will be 1 (uniform distribution between two discrete cases).
-Since $\iota(A)$ will be the maximally mixed state, which also corresponds to entropy 1.
+Since we want the entropy for a single pure state to be zero, the entropy for case A will be 1 (uniform distribution between two discrete cases).
+Since $\iota(A)$ will be the maximally mixed state, this also corresponds to entropy 1.
 So far so good: $\iota$ maps entropy 1 to entropy 1!
@@
 The entropy of C is therefore infinite with respect to B which is infinite with respect to A.
-On the other hand, A, B and C will all correspond to the same maximally mixed state, therefore $\iota$ collapses together epistemic states at different level of infinity.
+On the other hand, A, B and C will all correspond to the same maximally mixed state, therefore $\iota$ collapses together epistemic states at different levels of infinity.
 Note that, in this case, $\iota$ picks the minimum entropy between the equivalent epistemic states.
 This is inconsistent with the idea that $E_{Q}$ represents states where knowledge is lost: that would pick the maximum possible entropy.
+
+\begin{figure}
+\includegraphics[scale=.4]{fig2}
+\caption{\footnotesize{}}\label{fig_map}
+\end{figure}
 
@@
 However, $\iota(D)$ is not the maximally mixed state and therefore the entropy is less than 1.
-This means that epistemic states that are isoentropic in $E_{HS}$ are not isoentropic $E_{Q}$.
+This means that epistemic states that are isoentropic in $E_{HS}$ are not isoentropic in $E_{Q}$.
 If we think as $E_{HS}$ and $E_{Q}$ as partially ordered by information entropy, then $\iota$ is not an order preserving map.
@@
 
-The seemingly innocuous assumption that $P(\lambda|P)$ is well defined for all preparations $P$ and for all states $\lambda$ imposes a classical probability space which follows the rules of classical information theory, and prevents the state space $\Lambda$ to support an epistemic structure compatible with quantum theory.
+The seemingly innocuous assumption that $p(\lambda|P)$ is well defined for all preparations $P$ and for all states $\lambda$ imposes a classical probability space which follows the rules of classical information theory, and prevents the state space $\Lambda$ from supporting an epistemic structure compatible with quantum theory.
 This suggests an obvious follow up: what if we start by assuming that the set of pure states $\Lambda$ supports the correct epistemic structure, the one associated with density matrices, and use the same definitions of $\psi$-ontic and $\psi$-epistemic?
 
-The difficulty here is that support and overlaps are not a well defined concept for density matrices.
-Yet, there are two different properties that non-overlapping functions have in classical probability theory.
-Two distribution $\rho_1$ and $\rho_2$ are non-overlapping if and only if:
-\begin{enumerate}
-	\item their product $\rho_1 \rho_2 = 0$ is zero
-	\item the probability of being in the support of the second given the first is zero.
-\end{enumerate}
-In quantum mechanics, if we consider two pure states $\psi_1$ and $\psi_2$, they lead to the following conditions:
-\begin{enumerate}
-	\item $|\psi_1><\psi_1|\psi_2><\psi_2| = 0$
-	\item $<\psi_2|\psi_1><\psi_1|\psi_2> = 0$.
-\end{enumerate}
-Both of these are satisfied if and only if $<\psi_1|\psi_2>=0$: the states are non-overlapping if and only if they are orthogonal.
-This is, not suprisingly, exactly what happens in classical probability: two distributions are non-overlapping if and only if they are orthogonal with respect to the inner product $\int \rho_1 \rho_2 d\lambda$.\footnote{We could also define the overlap in terms of the entropy of the mixture and we would find the same result.}
+We can rephrase condition \ref{ontic_condition} directly in terms of density matrices as $\rho_1 \rho_2 = 0$ which, for pure states, becomes:
+\begin{equation}
+	|\psi_1><\psi_1|\psi_2><\psi_2| = 0.
+\end{equation}
+That is, two pure states are non-overlapping if and only if $<\psi_1|\psi_2>=0$, i.e.
+if they are orthogonal.
+Alternatively, we could ask that the probability of being in the support of the second given the first be zero.
+For pure states, we get $<\psi_2|\psi_1><\psi_1|\psi_2> = 0$, which leads to the same condition.
+Or we could define the overlap in terms of the entropy of the mixture and still find the same result.
 
-Therefore we have the following result: if we want a set of pure states that are able to support the epistemic structure of quantum mechanics, they must be ``overlapping'' and therefore they cannot be take to be ontic.
-The reader is no less entitled to conclude that quantum mechanics does not allow $\psi$-ontic interpretations from this result than he would be entitled to conclude that quantum mechanics does not allow $\psi$-epistemic interpretations from PBR.
-He would also be entitled to conclude that the categorization is flawed.
+However we look at it, we reach the following result:
+\begin{align}\label{conclusion}
+	\parbox{2.8in}{Non-orthogonal pure quantum states overlap and are therefore epistemic.}
+\end{align}
+This should not be surprising: even in classical probability two distributions are non-overlapping if and only if they are orthogonal with respect to the inner product $\int \rho_1 \rho_2 d\lambda$.
+
+Therefore we have the following result: if we want a set of pure states that is able to support the epistemic structure of quantum mechanics, they must be ``overlapping'' and therefore we must have a $\psi$-epistemic model.
+This may seem in contradiction with PBR, which shows that no $\psi$-epistemic model exists.
+We believe it is not in contradiction, but rather that there is no model in the HS classification that is compatible with quantum mechanics.
+In other words, the categorization itself is flawed.
 
@@
 
-Let us now step back a moment. Why does assuming $p(\lambda|P)$ lead to an epistemic structure incompatible with the quantum one?
+Let us now step back a moment. Why does assuming $p(\lambda|P)$ is well defined lead to an epistemic structure incompatible with the quantum one?
 
-It is widely understood that one feature of quantum mechanics is non-contextuality: a measurement cannot be understood as revealing pre-existing values.
+It is widely understood that one feature of quantum mechanics is contextuality: a measurement cannot be understood as revealing pre-existing values.
 We may be tempted to think that we have one context before and one after the measurement.
-What we are seeing that this does not work because we a dual effect on preparations.
-As we cannot understand quantum measurements as selecting from a single joint probability distribution, we cannot understand quantum preparations as constructing a single joint probability distribution.
-To fully understand how this works, let us review how how probability spaces work in both classical and quantum mechanics.
+This does not work because we have a related dual effect on preparations.
+In the same way that we cannot understand quantum measurements as selecting from a single joint probability distribution for all properties, we cannot understand quantum preparations as constructing a single joint probability distribution over all properties.
+To fully understand how this works, let us review how probability spaces work in both classical and quantum mechanics.
 
@@
 In quantum mechanics we start with a similar structure: a set of states $\mathcal{H}$ which represents all possible cases (i.e.
-the Hilbert space) and a $\sigma$-algebra $\Sigma_{\mathcal{H}}$ that represents all statements of interests.
+the Hilbert space) and a $\sigma$-algebra $\Sigma_{\mathcal{H}}$ that represents all statements of interest.
 This is where the similarity ends.
@@
 And this seems to work in some cases.
-Given any density matrix $\rho$, there exist a (non-unique) privileged context: this is the one in which we can express $\rho = \sum p_i |\psi_i > < \psi_i|$ as the mix of orthogonal pure states.
+Given any density matrix $\rho$, there exists a (non-unique) privileged context: this is the one in which we can express $\rho = \sum p_i |\psi_i > < \psi_i|$ as the mix of orthogonal pure states.
 If $\rho$ is itself a pure state, any basis that includes it will be a privileged context.
@@
 
-Suppose we start with two density matrixes $\rho_1$ and $\rho_2$ with different eigenstates.
+Suppose we start with two density matrices $\rho_1$ and $\rho_2$ with different eigenstates.
 We then create the mixture $\rho = p_1 \rho_1 + p_2 \rho_2$.
-The contexts of the mixture will not be the same as the original contexts.
+The context of the mixture will not be the same as the original contexts.
 It will, in fact, even depend on $p_1$, $p_2$.
-That is, when mixing preparations we are not simply combining two probability distribution over the same context, like in classical mechanics, but we are combining the contexts themselves.
-Non-contextuality does not affect only measurement, but, dually, preparations as well.
-And this affect how information entropy is calculated and the overall epistemic structure.
-
-% TODO: mention tomography?
+That is, when mixing preparations we are not simply combining two probability distributions over the same context, like in classical mechanics, but we are combining the contexts themselves to create a new one.
+Contextuality not only affects measurements, but, dually, preparations as well.
+And this affects how information entropy is calculated and the overall epistemic structure.
 
@@
 This is exactly assuming non-contextuality.
-While this can be made to work for pure state to pure state transitions during measurement, mixing preparations leads to an epistemic structure that is fundamentally non-contextual, leading to inconsistencies with quantum mechanics.
+While this can be made to work for pure state to pure state transitions during measurement, mixing preparations leads to an epistemic structure that is fundamentally non-contextual, leading to inconsistencies with quantum information.
 To allow for contextuality we should assume that for each observable $O$ we have a probability $p_O(\lambda_O|P)$ of having prepared one of the eigenstates, such that $\sum_{\lambda_O} p_O(\lambda_O|P) = 1$.
+While we have a single state space $\Lambda$, we do not have a single probability space over the whole space.
 The entropy of the preparation would be the lowest entropy across all contexts.

### 2022-01-10 TeachingMaterial/ActionPrinciple/ActionPrinciple.tex
ActionPrinciple: First rough draft

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{assumptionsofphysics}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\newcommand\partitle[1]{\textsc{#1}.}

\begin{document}

\title{Geometrical and physical interpretation of the action principle}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
TBD
\end{abstract}

\maketitle

\section{Introduction}

The action principle is of paramount importance in physics, yet it is somewhat mysterious. It is, in essence, a geometrical relationship yet it is not clear where and why that relationship originates. The fact that the Lagrangian for a system is not unique, makes it even more difficult to understand what is the physical content it provides.

In this paper we give a full characterization that is both geometrical and physical. We will see that the only physical assumption needed to recover both Hamiltonian mechanics and the action principle is that of deterministic and reversible evolution. We will also see how the action principle arises from core property of divergence-free fields and closed two-forms. The action principle is essentially a general mathematical property of vector calculus and differential geometry.

We will first address the case of a single degree of freedom using standard vector calculus. This provides already all the geometrical and physical insights in a language that is accessible to most physicist. Then we proceed to the case of multiple independent degrees of freedom using differential geometry. These new tools, though unfamiliar to many physicists and engineers, are unfortunately needed. Yet, we will give it a treatment and a notation that hopefully will make them more intuitive.

\section{One degree of freedom}

We are going to present the simple case of one degree of freedom using the standard tools provided by vector calculus. It is very fortunate that this can be done at all, since it allows the majority of people with a physics or engineering background to grasp all the major points.

The first thing we need to do is recast the equations of motion in a form that will make apparent both their geometrical interpretation and their physical meaning. The setting is phase space extended with the time variable. That is, the space charted by position $q$, momentum $p$ and time $t$. It is precisely because we have only three variable that make the standard tools of vector calculus appropriate.

TODO: figure with a few trajectories in extended phase space

Once we have recast the equation of motion on the extended phase space, we can see how the action principle comes from a straightforward application calculus stemming from a property of divergenceless fields.

\subsection{Hamiltonian reformulation}

Let us quickly review the equations for a Hamiltonian system with a single degree of freedom in the extended phase space. We can chart all states at all times by using three variables, $\{q,p,t\}$. Given a set of initial conditions $\{q_0,p_0,t_0\}$, we can find a single evolution $\gamma = \{q(t), p(t), t\}$ by solving Hamilton's equations:
\begin{equation}\label{Hamilton_equations}
\begin{aligned}
	\frac{dq}{dt} &= \frac{\partial H}{\partial p} \\
	\frac{dp}{dt} &= -\frac{\partial H}{\partial q}.
\end{aligned}
\end{equation}
How can we better understand the physics and geometry captured by these equations?

Let us step back, and forget we have a Hamiltonian system. Suppose we simply have a system described by a single degree of freedom and assume the following:
\begin{align}\label{Detrev_assumptions}
	\parbox{2.8in}{The system undergoes deterministic and reversible evolution.}
\end{align}

Under this assumption, given a state at a moment in time, there is a well defined change in all variable which is described by the displacement vector field
\begin{equation}
	\vec{S} = \left\{ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right\}.
\end{equation}
If we consider a finite region in the extended phase space, given that motion is deterministic and reversible we expect the number of states flowing in and flowing out of the region to be the same. Equivalently, we expect the evolutions to neither converge or diverge. Mathematically, the displacement field must be divergence-free, that is
\begin{equation}
	\nabla \cdot \vec{S} = 0,
\end{equation}
to properly represent a deterministic and reversible evolution.

Since $\vec{S}$ is divergence-free, we can find a vector potential
\begin{equation}
	\vec{\theta} = \{\theta_q, \theta_p, \theta_t\}
\end{equation}
such that
\begin{equation}\label{displacement_has_potential}
	\vec{S} = -curl(\vec{\theta}).
\end{equation}
The minus sign is introduced to match conventions of differential geometry. Mathematically, this is analogous to what is done for a magnetic field or for an incompressible fluid, except we are look at the flow of states in space, momentum and time.

Given that $\vec{\theta}$ is a vector potential, we have the usual gauge arbitrariness since $curl(\vec{\theta} + grad(f)) = curl(\vec{\theta})$ and therefore the displacement field $\vec{S}$ remains unchanged. We therefore can choose so that we can set
\begin{equation}
	\vec{\theta} = \{\theta_q, 0, \theta_t\}
\end{equation}
without loss of generality.

The displacement field has a further constraint, since $S_t = dt/ dt = 1$. Given \ref{displacement_has_potential}, we have
\begin{align*}
	S_t &= - \left(\frac{\partial}{\partial q}  \theta_p - \frac{\partial}{\partial p}  \theta_q\right) \\
	&= - \left(\frac{\partial}{\partial q}  0 - \frac{\partial}{\partial p}  \theta_q\right) \\
	& = \frac{\partial \theta_q}{\partial p} = 1
\end{align*}
Integrating, we have $\theta_q = p + g(q,t)$ where $g(q,t)$ is an arbitrary function, which we can set to zero. Therefore we have:
\begin{equation}
	\vec{\theta} = \{p, 0, \theta_t\}
\end{equation}
Lastly, we rename the last component as $\theta_t = -H$ which leads to 
\begin{equation}\label{potential_final}
	\vec{\theta} = \{p, 0, -H\}.
\end{equation}

If we now expand each component of \ref{displacement_has_potential} using \ref{potential_final}, we have:\footnote{Note that $\frac{dp}{dt}$ is not the same as $\frac{\partial p}{\partial t}$. In the first case we are taking a total derivative along the evolution, and therefore the momentum can change. In the second case we are taking a partial derivative which is taken at constant $q$ and $p$ by definition, and therefore $\frac{\partial p}{\partial t}=0$.}
\begin{align*}
	S_q = \frac{dq}{dt}
	&= - \left( \frac{\partial}{\partial p} \theta_t - \frac{\partial}{\partial t} \theta_p \right) \\
	&= - \left( \frac{\partial}{\partial p} (-H) - \frac{\partial}{\partial t} 0 \right) \\
	& = \frac{\partial H}{\partial q}
\end{align*}
\begin{align*}
	S_p = \frac{dp}{dt}
	&= - \left( \frac{\partial}{\partial t} \theta_q - \frac{\partial}{\partial q} \theta_t \right) \\
	&= - \left( \frac{\partial}{\partial t} p - \frac{\partial}{\partial q} (-H) \right) \\
	& = - \frac{\partial H}{\partial q}
\end{align*}
\begin{align*}
	S_t = \frac{dt}{dt}
	&= - \left( \frac{\partial}{\partial q} \theta_p - \frac{\partial}{\partial p} \theta_q \right) \\
	&= - \left( \frac{\partial}{\partial q} 0 - \frac{\partial}{\partial p} p \right) \\
	& = 1
\end{align*}
We have recovered Hamilton's equations \ref{Hamilton_equations} as the equations for a deterministic and reversible system. The only thing that the equations say is that states move in time at the same rate (i.e. $\frac{dt}{dt} = 1$) with an incompressible flow (i.e. no states are created or destroyed). That is the whole physical and geometrical content of those equations.

\subsection{Action principle}

Let us now review the principle of least action. We chart all states at all times by using $\{q, \frac{dq}{dt}, t\}$. We have a Langrangian $L(q, \frac{dq}{dt}, t)$ and, given a path $\gamma = \{q(t), \frac{dq}{dt}(t), t\}$ with $t_1 \leq t \leq t_2$, we define the action as $\mathcal{A}[\gamma] = \int_{t_1}^{t_2} L(q, \frac{dq}{dt}, t) dt$. An actual evolution of the system is such that the action is stationary:
\begin{equation}
	\frac{\delta \mathcal{A}}{\delta \gamma} = 0.
\end{equation}
The connection between the Lagrangian and Hamiltonian formulation is given by
\begin{equation}
	L = p \frac{dq}{dt} - H.
\end{equation}
How can we connect this with the previous picture?

Let us step back, forget about Lagrangians, and resume the discussion from before. Let $\gamma$ be an evolution (i.e. a solution of the equations of motion) with endpoints $A$ and $B$. Consider the line integral of the vector potential $\vec{\theta}$ along $\gamma$. We have:
\begin{equation}
\begin{aligned}
	\int_A^B \vec{\theta} \cdot d\gamma &= \int^{t_1}_{t_0} \vec{\theta} \cdot \vec{S} dt \\
	&= \int^{t_1}_{t_0} \left(\theta_q \frac{dq}{dt} + \theta_p \frac{dp}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
	&= \int^{t_1}_{t_0} \left(p \frac{dq}{dt} + 0 \frac{dp}{dt} - H \frac{dt}{dt}\right) dt \\
	&= \int^{t_1}_{t_0} \left(p \frac{dq}{dt} - H\right) dt
\end{aligned}
\end{equation}
In other words, the integral of the vector potential over an evolution coincides with the action, and the Lagrangian is simply
\begin{equation}
	L = \vec{\theta} \cdot \vec{S}.
\end{equation}

Now let us vary the path and consider the variation of the line integral $\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}$. The original path $\gamma$ together with its variation $\gamma'$ form a closed line $\lambda$. We can then use Stoke's theorem to transform the line integral of $\vec{\theta}$ to a surface integral of its curl. That is:
\begin{align*}
	\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} = 
	&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma'} \\
	&= \oint_{\lambda} \vec{\theta}  \cdot d\lambda \\
	&= \iint_{\Sigma} curl(\vec{\theta}) \cdot d\vec{\Sigma} \\
	&= - \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma}.
\end{align*}
Since $\gamma$ is an evolution for the system, it is always tangent to $\vec{S}$. This means that the surface $\Sigma$ enclosed by $\gamma$ and its infinitesimal variation is also parallel to $\vec{S}$ and therefore $\vec{S} \cdot d\vec{\Sigma} = 0$. Which means
\begin{equation}
	\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} = 0.
\end{equation}
Which recover the action principle.

The action principle, then, is a direct consequence of the divergence-free nature of the displacement field $\vec{S}$. Every evolution $\gamma$ is a field line of $\vec{S}$. Any infinitesimal surface $\Sigma$ enclosed by an evolution and its variation, then, will also be tangent to the displacement field $\vec{S}$ and therefore the surface integral of $\vec{S}$ over $\Sigma$ will be zero. Given that $\vec{S}$ is divergence-free, the integral over $\Sigma$ can be expressed as the line integral of the vector potential $\vec{\theta}$ over the contour of $\Sigma$.

\subsection{Discussion}

Before proceeding to the more general case, there are a few things of note.

First of all, we should not that the real and only physical content of the discussion lies in condition \ref{Detrev_assumptions} that assumes deterministic and reversible motion. The displacement field $\vec{S}$ and its properties are the only mathematical entities that are directly physically meaningful. The vector potential $\vec{\theta}$ exists provided condition \ref{Detrev_assumptions} and its expression is not unique.

Since the Lagrangian is the scalar product between the displacement and its vector potential, it suffers all the problems of the vector potential: it is not unique and its value is not directly physically meaningful. The same is true for the action itself. While the Lagrangian and the action are clearly useful tools in physics, this result warrants caution in relying exclusively on them to understand the physical relevance and meaning of a theory.

The second point is that our geometrical and physical interpretation works for all Hamiltonians. Technically, not all Hamiltonian systems admit a Lagrangian, as one needs an invertible map between conjugate momentum $p$ and the velocity $\frac{dq}{dt}$. For example, the following Hamiltonian
\begin{equation}
	H = c |p|
\end{equation}
is the one for a free photon treated as a particle. This gives us
\begin{equation}
	\frac{dq}{dt}= c \frac{p}{|p|}
\end{equation}
which means the velocity allows us to recover the direction of the conjugate momentum, but not its norm (i.e. the trajectory of a photon is not enough to recover its full state). In these cases, the Lagrangian approach cannot be applied because $\vec{\theta} \cdot \vec{S}$ cannot be expressed as a function of position and velocity. Yet, if we express $L(q,p,t)$ in terms of conjugate momentum, the modified version of the variational principle still works.

Note that the failure can happen in the reverse direction too: there are Lagrangian systems for which we cannot express the Hamiltonian in terms of conjugate momentum. However, these are exactly the cases where the action principle fails to yield a single solution. Therefore we have the following points:
\begin{enumerate}
	\item the only Lagrangian systems in which the action principle yield a single solutions are the ones that admit a Hamiltonian formulation;
	\item all Hamiltonian systems admit an action principle formulation, even though that do not admit a Lagrangian.
\end{enumerate}

In our view, this is enough evidence to conclude that the Hamiltonian setting in the extended phase space is the most natural to understand the action principle.

\section{Multiple degree of freedom}

All our previous points will hold true in the generalization to multiple degrees of freedom. The only additional key insight is that determinism and reversibility will apply to each independent degree of freedom, meaning that we do not only preserve the total number of states of the system, but we preserve the fact that the total number of states is the product of the configurations given by each independent degree of freedom.

To express this requirement, the tools provided by differential calculus are not enough. We need elements of differential geometry, particularly symplectic geometry and contact manifolds, which is problematic for two reasons. First, these are not widely used (or understood) among physicists, and are not accessible to physics students. Second, the treatment of the subject is typically too abstract and uses notation that we find, in some cases, misleading to a physicist. This is unfortunate because the underlying concepts are quite accessible and provide a more general and robust foundations for many areas of physics, not just classical particle mechanics.

Given that we want to make the generalization understandable to a wide array of people with physical background, we are going to use enough of the standard terminology from differential geometry that our methods are recognizable to someone proficient in it, but we are going to use a notation that parallel Einstein's notation, so that it feels more familiar to a physicist.\footnote{The main differences are as follows. The vector basis will be noted as $e_a$ instead of $\frac{\partial}{\partial x_a}$. The covector basis will be noted as $e^a$ instead of $dx^a$. The exterior derivative will be noted as $\partial$ (e.g. $\partial \omega$) instead of $d$ (e.g. $d \omega$).}

\subsection{Hamiltonian reformulation}

Our system will be now composed of $N$ independent degrees of freedom. This means that we can chart all states at all times using $2N + $ variables $\xi^a = \{ q^i, p_i, t\}$. We will use $\xi^a$ when we want to span all variables; $q^i$ and $p_i$ when we want to span the position or the momentum of all degrees of freedom.

As before, we will have a displacement field
\begin{equation}
	\vec{S} = \frac{d\xi^a}{dt} e_{a} = \frac{dq^i}{dt} e_{q^i} + \frac{dp_i}{dt} e_{p_i} + \frac{dt}{dt} e_t.
\end{equation}
However, in this cases we will not use this object to fully characterize the geometry of the space.

The main idea is that we wont be able to quantify the number of states identified by each degree of freedom. As degrees of freedom are bi-dimensional, this means quantifying the areas of two-dimensional surfaces. Therefore we introduce the two-form
\begin{equation}
	\omega = \omega_{ab} \, e^a \otimes e^b,
\end{equation}
which we call the state counting form. The idea is that a pair of two vectors $\vec{v}$ and $\vec{w}$ will identify a parallelogram and 
\begin{equation}
	\omega(\vec{v}, \vec{w}) = \omega_{ab} v^a w^b
\end{equation}
quantifies the number of states over its surface. One can think of $\omega$ as a rank 2 tensor: a linear function of two vectors.

The form $\omega$ will need to satisfy additional condition to be physically meaningful. First of all, it will need to be anti-symmetric. The parallelogram identified by $\vec{v}$ and $\vec{w}$ in that order will be the same as the one identified by $\vec{w}$ and $\vec{v}$ therefore the area is the same. However, the switch changes handedness, and therefore will introduce a minus sign. Therefore we have:
\begin{equation}
	\omega(\vec{v}, \vec{w}) = - \omega(\vec{w}, \vec{v}).
\end{equation}
This will mean the form will have $N(2N+1) = \sum_1^{2N}i$ independent components.

The form $\omega$ will also be closed, meaning
\begin{equation}\label{mdof_closed_form}
	\oiint \omega = 0
\end{equation}
over all closed surfaces. In fact, consider a parallelopiped. Given that the degrees of freedom are independent, and we have deterministic and reversible motion, two opposite sides will contain the same number of states. Given that the two opposite sides will have opposite orientation, their total contribution will be zero. This logic can be extended to all three-dimensional differential volumes, as they can be decomposed into infinitesimal parallelopipeds.

Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative of a covector $\theta$. We have
\begin{equation}
\begin{aligned}
	\theta &= \theta_a e^a = \theta_{q^i} e^{q^i} + \theta_{p_i} e^{p_i} + \theta_t e^t \\
	\omega &= - \partial \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
\end{aligned}
\end{equation}

In the generalization we need to distinguish between vectors, like $\vec{S}$, covectors, like $\theta$, and two-forms, like $\omega$. In three dimension, to each parallelogram there is only one perpendicular direction, which allows us to blur the line between the object in the following way. Each covector $\theta(\vec{S})$ can be expressed as a scalar product $\vec{\theta} \cdot \vec{S}$ with a suitable vector $\vec{\theta}$. Each two-form $\omega(\vec{v}, \vec{w})$ can be expressed as the triple product $\vec{\omega} \cdot \vec{v} \times \vec{w}$. The generalization of the curl is no longer an operation between vectors, but an operation that takes a one-form and returns a two-form.

Now we need to further characterize the $\omega$. Variables $q^i$ and $p_i$ form independent degrees of freedom. This means that only surfaces spanned by matching position and momentum will define actual states, while all other combination will define no state. Therefore:
\begin{equation}\label{canonical_conditions}
\begin{aligned}
	\omega(e^{q^i}, e^{p_j}) = \omega_{q^i p_j} &= \delta^i_j = - \omega_{p_j q^i} \\
	\omega(e^{q^i}, e^{q^j}) = \omega_{q^i q^j} &= 0 \\
	\omega(e^{p_i}, e^{p_j}) = \omega_{p_i p_j} &= 0
\end{aligned}
\end{equation}
The first condition sets $N^2$ independent components while the last two set $N(N-1)/2$ independent components each. This totals $N(2N-1) = \sum_1^{2N-1}i$ independent components.

Assuming determinism and reversibility, time evolution contributes no new states, it just moves them in time. That is, no states appear or disappear from any degree of freedom. The displacement field doesn't define any state, no matter which other vector is paired with. Mathematically, we say that the displacement kills the form
\begin{align}\label{mdof_displacement_kills}
	\omega(\vec{S}, \cdot) = 0.
\end{align}
This sets another $2N$ components (since $\omega(\vec{S}, \vec{S}) = 0$ is already set by anti-symmetry). All independent components are now set.

Those familiar with symplectic and contact geometry, will recognize that conditions \ref{canonical_conditions} mean that $q^i$ and $p_i$ are canonical coordinates and, by Darboux's theorem, we will be able to write the covector potential as
\begin{equation}\label{mdof_potential_expression}
	\theta = p_i e^{q^i} - H e^t.
\end{equation}
Let us convince ourselves of this result by constructing a proof that is similar to what we have done for the single degree of freedom.

First, we express \ref{canonical_conditions} in terms of the covector potential. We have:
\begin{equation}\label{canonical_potential_conditions}
\begin{aligned}
	\omega(e^{q^i}, e^{p_j}) &= (-d\theta)_{q^i p_j} = -(\partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i}) = \delta^i_j \\
	\omega_{q^i q^j} &= (-d\theta)_{q^i q^j} = -(\partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i}) = 0 \\
	\omega(e^{p_i}, e^{p_j}) &= (-d\theta)_{p_i p_j} = -(\partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i}) = 0
\end{aligned}
\end{equation}
We can use our gauge freedom to set $\theta_{p_1} = 0$, much in the same way we did for the simpler case. We now have $\partial_{q^1} \theta_{p_1} = 0$ and, by the first condition, $\partial_{p_1} \theta_{q^1} = 1$. Integrating, we have $\theta_{q^1} = p_1 + g(q^i, p_2, p_3, ..., t)$ where $g$ is an arbitrary function which we can set to zero. Therefore we have:
\begin{equation}
	\theta = p_1 e^{q^1} + 0 e^{p_1} + \theta_{q^2} e^{q^2} + \theta_{p_2} e^{p_2} + ... + \theta_{t} e^{t}.
\end{equation}

Note that the components for the first degree of freedom do not depend on the other degrees of freedom. That is, for all $i>1$, $\partial_{q^i} \theta_{q^1} = \partial_{p_i} \theta_{q^1} = \partial_{q^i} \theta_{p_1} = \partial_{p_i} \theta_{p_1} = 0$. But by using conditions \ref{canonical_potential_conditions}, we find that the converse is true as well: the components of all other degrees of freedom do not depend on the first. That is, for all $i>1$, $\partial_{q^1} \theta_{q^i} = \partial_{p_1} \theta_{q^i} = \partial_{q^1} \theta_{p_i} = \partial_{p_1} \theta_{p_i} = 0$.

We can then use, again, our gauge freedom with a function that does not depend on the first two variables. And we can use to set $\theta_{p_2} = 0$. And, with the same reasoning, we will be able to set $\theta_{q^2} = p_2$. And then, again, find that the first two degrees of freedom do not depend on the others, etc. At the end, we will find \ref{mdof_potential_expression}.

At this point, simply calculate $d\theta(S, \cdot ) $. The components of $d\theta$ are:
\begin{equation*}
	(d\theta)_{ab} = \begin{bmatrix}
		0 & \delta^j_i & - \partial_{q^i} H \\
		-\delta^i_j & 0 & - \partial_{p_i} H \\
		\partial_{q^j} H & \partial_{p_j} H & 0
	\end{bmatrix}
\end{equation*}
We have
\begin{align*}
	d\theta(S, \cdot )  &= S^a (d\theta)_{ab} e^b = 0 \\
	&= (S^{q^i}(d\theta)_{q^ib} + S^{p_i}(d\theta)_{p_ib} + S^{t}(d\theta)_{tb}) e^b \\
	&= (S^{q^i}(d\theta)_{q^iq^j} + S^{p_i}(d\theta)_{p_iq^j} + S^{t}(d\theta)_{tq^j}) e^{q^j} + \\
	& (S^{q^i}(d\theta)_{q^ip_j} +  S^{p_i}(d\theta)_{p_ip_j} + S^{t}(d\theta)_{tp_j}) e^{p_j} + \\
	& (S^{q^i}(d\theta)_{q^it} + S^{p_i}(d\theta)_{p_it} + S^{t}(d\theta)_{tt}) e^t \\
	&= (-S^{p_i}\delta^i_j + S^{t}\partial_{q^j} H ) e^{q^j} + \\
	& (S^{q^i}\delta^j_i +  S^{t}\partial_{p_j} H) e^{p_j} + \\
	& (-S^{q^i} \partial_{q^i} H - S^{p_i} \partial_{p_i} H) e^t \\
\end{align*}

All components must be zero, therefore we have the following three equations:
\begin{align*}
	S^{p_j} &= S^{t} \partial_{q^j} H \\
	S^{q^j} &= - S^{t}\partial_{p_j} H \\
	-S^{q^i} \partial_{q^i} H - S^{p_i} \partial_{p_i} H &= S^{t}\partial_{p_i} H \partial_{q^i} H - S^{t} \partial_{q^i} H \partial_{p_i} H = 0
\end{align*}
Note that the last expression is not a new equation: it is identical to zero given the previous two equations. Since $S^t = 1$, we have:
\begin{align*}
	S^{p_j} &= \partial_{q^j} H \\
	S^{q^j} &= - \partial_{p_j} H
\end{align*}
Which recovers Hamilton's equation for multiple degrees of freedom.

Once again, we have found that the whole physical content is included in conditions \ref{Detrev_assumptions}. The geometry in the extended phase space is fully specified by the state counting form $\omega$. This does not give define lengths and angles, like the metric tensor does for a more standard geometric setting. It defines areas of two-dimensional surfaces in terms of states, and angles between two-dimensional surfaces in terms of the degree of independence between degrees of freedom. What the equations say, is simply that states move in time at the same rate (i.e. $\frac{dt}{dt} = 1$) with an deterministic and reversible flow (i.e. no states are created or destroyed on any independent degree of freedom).

\subsection{Action principle}
Having reformulated Hamiltonian mechanics in the extended phase space using the language of differential geometry, the geometrical interpretation of the action principle is essentially the same.

The Lagrangian is the potential covector applied to the displacement field. That is
\begin{equation}
\begin{aligned}
L &= \theta(\vec{S}) = \theta_a S^a \\
&= \theta_{q^i} S^{q^i} + \theta_{p_i} S^{p_i} + \theta_{t} S^{t} \\
&= p_i \frac{dq^i}{dt} + 0 \frac{dp_i}{dt} - H \frac{dt}{dt} \\
&= p_i \frac{dq^i}{dt} - H.
\end{aligned}
\end{equation}

As before, we will consider the variation of the line integral $\delta \int_{\gamma} \theta(d\gamma)$, which can be expressed as
\begin{equation}
	\begin{aligned}
		\int_A^B \theta(d\gamma) &= \int^{t_1}_{t_0} \theta(\vec{S}) dt \\
		&= \int^{t_1}_{t_0} \left(p_i \frac{dq^i}{dt} - H \right) dt
	\end{aligned}
\end{equation}
The original path $\gamma$ together with its variation $\gamma'$ form a closed line $\lambda$. We can then use Stoke's theorem to transform the line integral of $\vec{\theta}$ to a surface integral of its exterior derivative. That is:
\begin{align*}
	\delta \int_{\gamma} \theta(d\vec{\gamma}) = 
	&= \int_{\gamma} \theta(d\vec{\gamma}) - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma'} \\
	&= \oint_{\lambda} \theta(d\vec{\lambda}) \\
	&= \iint_{\Sigma} \partial \theta (\vec{\Sigma}) \\
	&= - \iint_{\Sigma} \omega(d\vec{\Sigma}).
\end{align*}
One side of each $d\vec{\Sigma}$ will be $d\gamma$, therefore $\omega(d\vec{\Sigma}) = \omega(d\vec{\gamma}, \vec{v}) = \omega(\vec{S}dt, d\lambda)$ where $d\lambda$ is the direction of the variation. Since \ref{mdof_displacement_kills}, we have
\begin{equation}
	\begin{aligned}
	\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} &= - \iint_{\Sigma} \omega(d\vec{\Sigma}) \\
	&= - \iint_{\Sigma} \omega(\vec{S}, d\lambda) dt \\
	&= - \iint_{\Sigma} 0 dt = 0
	\end{aligned}
\end{equation}

The action principle, then, is a direct consequence of the geometry set by \ref{mdof_closed_form} and \ref{mdof_displacement_kills}, which is the generalization of the divergence-free nature displacement field $\vec{S}$. Every evolution $\gamma$ is a field line of $\vec{S}$. Any infinitesimal surface $\Sigma$ enclosed by an evolution and its variation, then, will also be tangent to the displacement field $\vec{S}$ and therefore will contribute no states, $\omega(d\Sigma)$ will be zero. Given that $\omega$ is closed, the integral over $\Sigma$ can be expressed as the line integral of the covector potential $\theta$ over the contour of $\Sigma$.

\subsection{Discussion}

Apart from the technicalities needed to properly frame the problem, multiple degrees of freedom do not add much conceptually. All geometrical and physical content within the theory is about counting states and keeping track of how they move in time, which is done by the two-form $\omega$. As before, the assumption of deterministic and reversible evolution does all the work.

The additional key assumption is that the degrees of freedom must be independent. If the degrees of freedom were not independent, then the number of states within a four-dimensional parellopiped would not be given by the count of states on the sides and the angle. We would need additional structure. Conversely, if the degrees of freedom would not remain independent, the displacement field $\vec{S}$ would only preserve the count of states of volumes, not over each degree of freedom, and therefore $\omega(\vec{S}, \cdot) \neq 0$.

The comments on the lack of strict physicality for the values of the Lagrangian and the action remains unchanged. They both depend on a choice of gauge. The comment on the centrality of the extended phase space to understand the action principle is only reinforce. Additionally, note that by changing $t$ to $ct$ expression \ref{mdof_potential_expression} becomes
\begin{equation}\label{mdof_potential_relativistic}
	\theta = p_i e^{q^i} - \frac{H}{c} e^{ct}.
\end{equation}
which bears striking resemblance to the relativistic four-momentum. While it goes outside the scope of this article, the idea is that, in this setting, we already have some pre-relativistic features, even though we have not introduced a metric tensor.

\section{Conclusion}

We have given a geometrical interpretation of the action principle that is also fully physically motivated by the assumption of deterministic and reversible evolution. The key insight is that a path $\gamma$ and a variation form a closed curve which encloses a surface $\Sigma$. If the path is an actual evolution for the system, all states flow along the surface $\Sigma$, not through it. Using Stoke's theorem, the zero flow condition through the surface $\Sigma$ is transformed to a zero integral condition on its boundary, on the path $\gamma$ and its variation, leading to the principle of least action.

We have seen that both the Lagrangian and the action depend on the potential, which means they are gauge dependent. This makes the numerical values of those objects not strictly physical. We have also seen that the interpretation given also works for those Hamiltonian systems that do not have a corresponding Lagrangian, meaning that the most general expression for the action principle lies on the extended phase space formulation of classical mechanics. We have also seen that elements from symplectic and contact geometry are needed for the generalization and that they can be formulated in a language that is less abstract and closer to the physics.

All considered, we believe this way of looking at classical mechanics provides a more complete characterization and better insights than the usual treatments.

\bibliography{bibliography}

\end{document}

### 2022-01-10 TeachingMaterial/ActionPrinciple/ActionPrinciple.tex
ActionPrinciple: review of section 2

@@
 We will see that the only physical assumption needed to recover both Hamiltonian mechanics and the action principle is that of deterministic and reversible evolution.
-We will also see how the action principle arises from core property of divergence-free fields and closed two-forms.
-The action principle is essentially a general mathematical property of vector calculus and differential geometry.
+We will also see how the action principle arises as a feature of divergence-free fields and closed two-forms.
+In a nutshell, we will see that the action principle arises as a general mathematical property of vector calculus and differential geometry, its physical significance is of less importance than the assumption of determinism and reversibility it requires.
 
@@
 
-TODO: figure with a few trajectories in extended phase space
+FIGURE: a few trajectories in extended phase space
 
-Once we have recast the equation of motion on the extended phase space, we can see how the action principle comes from a straightforward application calculus stemming from a property of divergenceless fields.
+Once we have recast the equation of motion on the extended phase space, we can see how the action principle comes from a straightforward application of vector calculus in the case of divergence-free fields.
 
@@
 \begin{align}\label{Detrev_assumptions}
-	\parbox{2.8in}{The system undergoes deterministic and reversible evolution.}
+	\parbox{2.8in}{the system undergoes deterministic and reversible evolution.}
 \end{align}
 
-Under this assumption, given a state at a moment in time, there is a well defined change in all variable which is described by the displacement vector field
+Under this assumption, given a state at a moment in time, there is a well-defined change in all variable which is described by the displacement vector field
 \begin{equation}
@@
 \end{equation}
-If we consider a finite region in the extended phase space, given that motion is deterministic and reversible we expect the number of states flowing in and flowing out of the region to be the same.
+If we consider a finite region in the extended phase space, given that the evolution is deterministic and reversible we expect the number of states flowing in and flowing out of the region to be the same.
 Equivalently, we expect the evolutions to neither converge or diverge.
@@
 
+FIGURE: region in extended phase space with trajectories flowing through
+
 Since $\vec{S}$ is divergence-free, we can find a vector potential
@@
 \begin{equation}\label{displacement_has_potential}
-	\vec{S} = -curl(\vec{\theta}).
+	\vec{S} = - \nabla \times \vec{\theta}.
 \end{equation}
-The minus sign is introduced to match conventions of differential geometry.
-Mathematically, this is analogous to what is done for a magnetic field or for an incompressible fluid, except we are look at the flow of states in space, momentum and time.
+The minus sign is introduced to match conventions in differential geometry. Mathematically, this is analogous to what is done for a magnetic field or for an incompressible fluid.
 
-Given that $\vec{\theta}$ is a vector potential, we have the usual gauge arbitrariness since $curl(\vec{\theta} + grad(f)) = curl(\vec{\theta})$ and therefore the displacement field $\vec{S}$ remains unchanged.
-We therefore can choose so that we can set
+Given that $\vec{\theta}$ is a vector potential, we have the usual gauge arbitrariness since $\nabla \times(\vec{\theta} + \nabla f) = \nabla \times \vec{\theta}$ and therefore the displacement field $\vec{S}$ remains unchanged.
+By choosing $f$ appropriately, we can set
 \begin{equation}
@@
 	&= - \left( \frac{\partial}{\partial p} (-H) - \frac{\partial}{\partial t} 0 \right) \\
-	& = \frac{\partial H}{\partial q}
+	& = \frac{\partial H}{\partial p}
 \end{align*}
@@
 
-Now let us vary the path and consider the variation of the line integral $\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}$.
-The original path $\gamma$ together with its variation $\gamma'$ form a closed line $\lambda$.
-We can then use Stoke's theorem to transform the line integral of $\vec{\theta}$ to a surface integral of its curl.
+FIGURE: trajectory with variation forming a closed loop. Another small area with trajectories flowing through
+
+Now let us consider the variation of the line integral $\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}$ caused by an infinitesimal change in the path.
+The original path $\gamma$ together with its variation $\gamma'$ form a contour $\partial \Sigma$ which encloses the surface $\Sigma$.
+We can therefore use Stoke's theorem to transform the line integral of $\vec{\theta}$ over $\partial \Sigma$ to a surface integral of the curl $\nabla \times \vec{\theta}$ over $\Sigma$.
 That is:
@@
 	&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma'} \\
-	&= \oint_{\lambda} \vec{\theta}  \cdot d\lambda \\
-	&= \iint_{\Sigma} curl(\vec{\theta}) \cdot d\vec{\Sigma} \\
+	&= \oint_{\partial \Sigma} \vec{\theta}  \cdot d\lambda \\
+	&= \iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
 	&= - \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma}.
@@
 Given that $\vec{S}$ is divergence-free, the integral over $\Sigma$ can be expressed as the line integral of the vector potential $\vec{\theta}$ over the contour of $\Sigma$.
+Asking for a stationary action means asking for the flow of $\vec{S}$ to be zero across the surface given by the path and its variation, which happens only if the path is tangent to $\vec{S}$ and is therefore an evolution.
+In practice, it is a very roundabout way to ask that the path be tangent to the displacement vector.
 
@@
 
-Before proceeding to the more general case, there are a few things of note.
+Before proceeding to the more general case, there are two important things to note.
 
-First of all, we should not that the real and only physical content of the discussion lies in condition \ref{Detrev_assumptions} that assumes deterministic and reversible motion.
+First of all, the true (and only) physical content of the discussion lies in condition \ref{Detrev_assumptions} that assumes deterministic and reversible motion.
 The displacement field $\vec{S}$ and its properties are the only mathematical entities that are directly physically meaningful.
-The vector potential $\vec{\theta}$ exists provided condition \ref{Detrev_assumptions} and its expression is not unique.
+The vector potential $\vec{\theta}$ is not uniquely defined by condition \ref{Detrev_assumptions}, though it is extremely useful as it allows us to map the space of all possible displacement fields of interested with a single ``scalar'' function.
 

### 2022-01-16 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: introduction from Andrea with other changes

@@
-%\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
 \documentclass[twocolumn,prl,floatfix,superscriptaddress]{revtex4-2}
-
-\usepackage{amsthm, amsmath, amssymb, enumitem, graphicx, hyperref}
-\usepackage{graphicx, subfigure}
-\setlist{noitemsep}
-\hypersetup{
-	colorlinks=true,
-	citecolor=blue,
-	urlcolor=blue,
-	linkcolor=blue
-}
-\urlstyle{same}
-
-%\usepackage{graphicx, subfigure, amsfonts,amssymb,amsmath, array, gensymb,fontenc,color,lipsum}
+\usepackage{graphicx, subfigure, amsfonts, amssymb, amsmath, array, gensymb, fontenc, color, lipsum}
 %\documentclass[aps,twocolumn,showpacs,preprintnumbers]{revtex4}
@@
 %\usepackage{amssymb}
-%\usepackage{hyperref}
-%\hypersetup{colorlinks= true, allcolors=blue}
+\usepackage{hyperref}
+\hypersetup{colorlinks= true, allcolors=blue}
 %\setcitestyle{aysep={}}
@@
 
+\newcommand\partitle[1]{\textsc{#1}.}
 \begin{document}
@@
 \author{Andrea Oldofredi}
-\affiliation{Department of Philosophy, University of Lisbon, Portugal}
+\affiliation{Centre of Philosophy (CFUL), University of Lisbon, Portugal}
 \author{Christine Aidala}
@@
 
-%- Status of field and impact of HS
-%- Mention the PBR Theorem explicitly
-%- Claim that HS does not work for quantum epistemic cases, therefore invalidating all results
+%---e.g.\ standard Quantum Mechanics (QM) is classified as $\psi$-ontic and $\psi-$complete, hidden variable theories such as Bohm causal approach \cite{Bohm:1952aa} and Nelson mechanics \cite{Nelson:1966aa} are both considered $\psi$-ontic and $\psi$-incomplete, whereas Relational QM \cite{Rovelli:1996} and Quantum Bayesianism \cite{Fuchs:2002} are $\psi-$epistemic.
+%Since its publication HS paper has been subjected to several analyses and extensions as well as critical discussions showing some limitations of this approach (\cite{Leifer:2014, Leifer:2014b}, \cite{Leifer:2013}, \cite{Hermens:2021}, \cite{Wood:2015}, \cite{Ringbauer:2015}, \cite{Mazurek:2016}, \cite{Bartlett:2012}, \cite{Oldofredi:2020b}, \cite{Lewis:2012}, \cite{Ladyman:2021}).
+However, still today HS distinction is employed to argue what types of interpretations are possible.
+Referring to this, the PBR theorem \cite{PBR:2012}---a remarkable result derived on top of HS framework---rules out $\psi$-epistemic models.
+More precisely, PBR argued that in every model reproducing the statistics and predictions of QM the quantum state $\psi$ must represent real physical properties of the system under consideration---i.e.\ models must be $\psi-$ontic.
+Consequently, quantum theories cannot be $\psi-$epistemic.
+Given the importance of this claim and the resonance that such a theorem had (\cite{Renner:2012}, \cite{Hardy:2013}, \cite{Maroney:2014}, \cite{Patra:2013}, \cite{Mansfield:2016}, \cite{Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014}, \cite{Ben:2017}, \cite{Rizzi:2018}, \cite{Oldofredi:2021}, \cite{DeBrota:2019}), it is worthwhile to look at HS framework again with more scrutiny in order to understand whether it is a technically sound starting point to draw conclusions on the nature of the quantum state.
 
-TODO intro with status of the field
+In 2012 Pusey, Barrett and Rudolph published a formal result in \emph{Nature Physics}, widely known as the PBR theorem, showing that ``if the quantum state merely represents information about the real physical state of a system, then experimental predictions are obtained that contradict those of quantum theory'' (\cite{PBR:2012}, p.\ 475).
+Alternatively stated, PBR argued that in every model reproducing the statistics and predictions of Quantum Mechanics (QM) the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi-$ontic.
+Consequently, quantum theories cannot be $\psi-$epistemic.
+
+Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b}, \cite{Renner:2012}, \cite{Colbeck:2017}, \cite{Hardy:2013}, \cite{Maroney:2014}, \cite{Patra:2013}, \cite{Mansfield:2016}, \cite{Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014}, \cite{Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information.
+On the other, scholars recently shown that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017}, \cite{Rizzi:2018}, \cite{Oldofredi:2021}, \cite{DeBrota:2019}).
+
+In this essay we are not going to rebut the theorem itself.
+We instead aim to carefully re-examine one of its premises, i.e.\ Harrigan and Spekkens (HS) classification between $\psi-$ontic and $\psi-$epistemic ontological models  \cite{Harrigan:2010} on top of which the PBR result is derived.
+We believe that a careful examination of this assumption forces us to re-evaluate the significance of the PBR conclusion.
+
+HS proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in that case the model is called $\psi-$ontic, or to some observer information, making it $\psi-$epistemic.
+While the original aim of this classification was to clarify Einstein's view of quantum theory, HS framework has been widely employed in the literature not only to categorize different interpretations, but also to argue what types of interpretations are admissible (\cite{Leifer:2013}, \cite{Leifer:2017}, \cite{Branciard:2014} \cite{Hermens:2021}, \cite{Wood:2015}, \cite{Ringbauer:2015}, \cite{Mazurek:2016}, \cite{Bartlett:2012}; cf.\ \cite{Oldofredi:2020b}, \cite{Lewis:2012}, \cite{Ladyman:2021} for critical discussions).
+Given the influence that HS framework and the PBR theorem have in the quantum foundations, it is worthwhile to look again at the former with more scrutiny in order to understand whether it is a sound basis to draw conclusions on the nature of the quantum state.
 
@@
 \begin{enumerate}
-	\item the information entropy as calculated on epistemic states does not match the one provided by quantum mechanics, meaning that the two epistemic structures are incompatible;
-	\item the density matrices associated with pure states overlap for non-orthogonal states (i.e. $\rho_\psi \rho_\phi \neq 0$), meaning $\psi$-ontic models are ruled out.
+	\item the information entropy as calculated on epistemic states does not match the one provided by quantum mechanics, meaning that the epistemic structures provided by QM and HS are incompatible;
+	\item the density matrices associated with pure states overlap for non-orthogonal states (i.e.
+$\rho_\psi \rho_\phi \neq 0$), meaning that $\psi$-ontic models are ruled out, contrary to the PBR claim.
 \end{enumerate}
-Our conclusion is that the ontological model is fundamentally flawed.
-The problem lies in assuming that the epistemic state identifies a single probability distribution over the whole set of ontic states.
-We show that this, in essence, is an assumption of non-contextuality.
+Hence, our conclusion is that HS distinction is fundamentally flawed.
+The main problem lies in assuming that the epistemic state identifies a single probability distribution over the whole set of ontic states.
+We show that this, in essence, is an assumption of non-contextuality, which is also inherited by the PBR theorem, casting doubts on its validity.
 
@@
 
-We briefly review the main features of the classification provided by Harrigan and Spekkens (HS)\cite{Harrigan:2010}.
-We start with the usual operational setting for quantum mechanics.
+We briefly review the main features of the classification provided by HS starting with the usual operational setting for QM.
 We have a preparation protocol $P$, which will be associated with a density operator $\rho$ on the relevant Hilbert space, and a measurement protocol $M$, represented by a POVM $\{ E_k\}$ where each $k$ represents a possible measurement outcome.
@@
 
-An ontological model, as defined by HS, additionally assumes that there exists a set of states $\Lambda$, called \textbf{ontic states}, that provide the complete specification of a given physical object.
+An ontological model, as defined by HS, additionally assumes that there exists a set of states $\Lambda$, called \textbf{ontic states}, that provide the complete specification of the properties of a given physical object.
 A preparation $P$ will prepare a particular ontic state $\lambda$ according to the probability distribution $p(\lambda | P)$, which is referred to as \textbf{epistemic state}.
-The measurement outcome will depend only on the ontic state with probability $p(k|\lambda, M)$.(\cite{Harrigan:2010}, p.\ 128).
+The measurement outcome will depend only on the ontic state with probability $p(k|\lambda, M)$ (\cite{Harrigan:2010}, p.\ 128).
 This leads to the following expression:
@@
 In a \textbf{$\psi$-ontic} model the wave function is a physical property of the ontic state, in the sense that given an ontic state $\lambda$ there is only one pure state preparation $P_\psi$ that could have prepared $\lambda$.
-As we can see in figure \ref{overlap}, this happens if the probability distributions do not overlap, i.e.
-if we have
+As we can see in figure \ref{overlap}, this happens if the probability distributions do not overlap, i.e.\ if we have
 \begin{equation}\label{ontic_condition}
@@
 for all pairs of states $\psi$ and $\phi$.
-If a model is not $\psi$-ontic, it is called \textbf{$\psi$-epistemic} and the wave function is taken to represent knowledge about the state preparation.
+If a model is not $\psi$-ontic, then it is \textbf{$\psi$-epistemic}.
+In this case, $\lambda$ can be described by more than one quantum state and the wave function is taken to represent knowledge about the state preparation---in such models quantum states generate overlapping probability distributions over $\Lambda$ as shown in figure \ref{overlap} (b).
 A $\psi$-ontic model is said \textbf{$\psi$-complete} if the quantum states and the ontic states coincide.
@@
 When we take the particle from the oven, we are sampling from a Boltzmann distribution over different energies.
-Unlike energy, temperature does not partition phase space because the same particle state could have been prepared by ovens at different temperature.
+However, unlike energy, temperature does not partition phase space because the same particle state could have been prepared by ovens at different temperature.
 Temperature is a property of the preparation and therefore an \emph{epistemic property}.
 
-While the original aim of this classification was to clarify Einstein's view of quantum theory, it has been used in the literature to classify different interpretations (e.g.
+%While the original aim of this classification was to clarify Einstein's view of quantum theory, it has been used in the literature to classify different interpretations (e.g.
 hidden variable theories such as Bohm pilot-wave theory \cite{Bohm:1952aa} and Nelson mechanics \cite{Nelson:1966aa} are both considered $\psi$-ontic and $\psi$-incomplete) and also to argue what types of interpretations are possible.
-Most notably, the PBR theorem rules out any $\psi$-epistemic model (TODO beef up this paragraph with citations).
+Most notably, the PBR theorem \cite{PBR:2012} rules out any $\psi$-epistemic model.
+More precisely, PBR argued that in every model reproducing the statistics and predictions of Quantum Mechanics (QM) the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi-$ontic.
+Consequently, quantum theories cannot be $\psi-$epistemic.(TODO beef up this paragraph with citations).
 It is therefore worthwhile to look at the ontological model itself with more scrutiny.
@@
 \includegraphics[scale=.7]{ontic}
-\caption{\footnotesize{Harrigan and Spekkens distinction between $\psi-$ontic (a) and $\psi-$epistemic (b) ontological models.
-TODO: change the notation to $p(\lambda | P_\psi)$; use product of probability equal to zero}}\label{overlap}
+\caption{\footnotesize{Harrigan and Spekkens distinction between $\psi-$ontic (a) and $\psi-$epistemic (b) ontological models.}}\label{overlap}
 \end{figure}
@@
 
-We look more closely at the space of epistemic states given by $p(\lambda | P)$ and equation \ref{epistemic_mixing}.
-To make the discussion more concrete, we will assume we have a $\psi$-complete ontological model of a single qubit representing the direction of spin.
+Let us consider more closely the space of epistemic states given by $p(\lambda | P)$ and equation \ref{epistemic_mixing}.
+To make the discussion more concrete, we assume to have a $\psi$-complete model of a single qubit representing the direction of spin.
 
-Suppose we have a protocol that prepares either z+ or z- with equal chance.
-This results in an epistemic state with equal distribution on z+ and z-.
-Now suppose we have another protocol that prepares x+ and x- again with equal chance.
-This time we have an epistemic state with equal distribution on x+ and x-.
-Note that the epistemic distributions do not overlap.
-Therefore, according to the HS categorization, these correspond to two ontologically distinct states, making the preparation axis an ontological property of the system, regardless of the orientation.
+Suppose we have a protocol that prepares either $z^+$ or $z^-$ with equal chance.
+This results in an epistemic state with equal distribution on $z^+$ and $z^-$.
+Similarly, suppose we have another protocol that prepares $x^+$ and $x^-$ with equal chance generating an epistemic state with equal distribution on $x^+$ and $x^-$.
+Note that such epistemic distributions do not overlap.
+Therefore, according to HS categorization, these correspond to two ontologically distinct states, making the preparation axis an ontological property of the system, regardless of the orientation.
 However, the respective density matrices would not just overlap: they would be the same maximally mixed state.
-Hence there seems to be a mismatch between the type of epistemic knowledge one can express in quantum mechanics through mixed states and the type of epistemic knowledge described in the HS model.
+Hence there seems to be a mismatch between the type of epistemic knowledge one can express in QM through mixed states and the type of epistemic knowledge described in the HS model.
 This needs to be investigated further.
@@
 In HS, the space of epistemic states $E_{HS}$ is the set of probability distributions over pure states, which is an infinite dimensional space.
-This is a direct consequence of defining the epistemic states as a distribution $p(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability to each ontological state $\lambda$.\footnote{We will follow the convention of not distinguishing between probability and probability density.} In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
+This is a direct consequence of defining the epistemic states as a distribution $p(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability to each ontological state $\lambda$ \footnote{We will follow the convention of not distinguishing between probability and probability density.}.
+In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
 Looking at the difference in dimension, the discrepancy is not minor.
@@
 For example, one may want to say that the axis of preparation is, in principle, an ontologically distinct property, but this is not experimentally distinguishable.
-Therefore distinct HS-epistemic states for each direction are projected by $\iota$ onto the same Q-epistemic state of maximum uncertainty.
-This sounds very plausible and appealing, but it turns out this is not at all the case.
-To show this we will concentrate on a single aspect of the map: how information entropy transforms under the map.
+Therefore, distinct HS-epistemic states for each direction are projected by $\iota$ onto the same Q-epistemic state of maximum uncertainty.
+This sounds very plausible and appealing, but it turns out this is not the case.
+To show this fact we will concentrate on a single aspect of the map: how information entropy transforms under $\iota$.
 While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is sufficient to show that the epistemic structures of $E_{HS}$ and $E_{Q}$ are different.
@@
 \begin{description}
-	\item[A] the poles, z+ and z-
-	\item[B] the equator, all states on the x/y planes
-	\item[C] the whole sphere, all points
+	\item[A] the poles, $z^+$ and $z^-$;
+	\item[B] the equator, all states on the x/y planes;
+	\item[C] the whole sphere, all points.
 \end{description}
-Since we want the entropy for a single pure state to be zero, the entropy for case A will be 1 (uniform distribution between two discrete cases).
+Since we want the entropy for a single pure state to be zero, the entropy for case \textbf{A} will be 1 (uniform distribution between two discrete cases).
 Since $\iota(A)$ will be the maximally mixed state, this also corresponds to entropy 1.
 So far so good: $\iota$ maps entropy 1 to entropy 1!
-However, note that A is a measure zero set of B, which is a measure zero set of C.
-The entropy of C is therefore infinite with respect to B which is infinite with respect to A.
-On the other hand, A, B and C will all correspond to the same maximally mixed state, therefore $\iota$ collapses together epistemic states at different levels of infinity.
+However, note that \textbf{A} is a measure zero set of \textbf{B}, which is a measure zero set of \textbf{C}.
+The entropy of \textbf{C} is therefore infinite with respect to \textbf{B} which is infinite with respect to \textbf{A}.
+On the other hand, \textbf{A}, \textbf{B} and \textbf{C} will all correspond to the same maximally mixed state, therefore $\iota$ collapses together epistemic states at different levels of infinity.
 Note that, in this case, $\iota$ picks the minimum entropy between the equivalent epistemic states.
@@
 \begin{description}
-	\item[D] the north pole and a point on the equator, z+ and x+.
+	\item[D] the north pole and a point on the equator, $z^+$ and $x^+$.
 \end{description}
@@
 If we think as $E_{HS}$ and $E_{Q}$ as partially ordered by information entropy, then $\iota$ is not an order preserving map.
-Therefore the epistemic structures represented by $E_{HS}$ and $E_{Q}$ are not isomorphic.
+Therefore, the epistemic structures represented by $E_{HS}$ and $E_{Q}$ are not isomorphic.
 
 The seemingly innocuous assumption that $p(\lambda|P)$ is well defined for all preparations $P$ and for all states $\lambda$ imposes a classical probability space which follows the rules of classical information theory, and prevents the state space $\Lambda$ from supporting an epistemic structure compatible with quantum theory.
+
 This suggests an obvious follow up: what if we start by assuming that the set of pure states $\Lambda$ supports the correct epistemic structure, the one associated with density matrices, and use the same definitions of $\psi$-ontic and $\psi$-epistemic?
-
 We can rephrase condition \ref{ontic_condition} directly in terms of density matrices as $\rho_1 \rho_2 = 0$ which, for pure states, becomes:
 \begin{equation}
-	|\psi_1><\psi_1|\psi_2><\psi_2| = 0.
+	|\psi_1\rangle\langle\psi_1|\psi_2\rangle\langle\psi_2| = 0.
 \end{equation}
-That is, two pure states are non-overlapping if and only if $<\psi_1|\psi_2>=0$, i.e.
+That is, two pure states are non-overlapping if and only if $\langle\psi_1|\psi_2\rangle=0$, i.e.
 if they are orthogonal.
 Alternatively, we could ask that the probability of being in the support of the second given the first be zero.
-For pure states, we get $<\psi_2|\psi_1><\psi_1|\psi_2> = 0$, which leads to the same condition.
+For pure states, we get $\langle\psi_2|\psi_1\rangle\langle\psi_1|\psi_2\rangle = 0$, which leads to the same condition.
 Or we could define the overlap in terms of the entropy of the mixture and still find the same result.
@@
 
-Therefore we have the following result: if we want a set of pure states that is able to support the epistemic structure of quantum mechanics, they must be ``overlapping'' and therefore we must have a $\psi$-epistemic model.
-This may seem in contradiction with PBR, which shows that no $\psi$-epistemic model exists.
-We believe it is not in contradiction, but rather that there is no model in the HS classification that is compatible with quantum mechanics.
+Therefore, we have the following result: if we want a set of pure states that is able to support the epistemic structure of QM, they must be ``overlapping'' and thus we must have a $\psi$-epistemic model.
+This may seem in contradiction with the PBR theorem.
+Nonetheless, we believe that there is no contradiction, since there is no model in the HS classification that is compatible with quantum theory.
 In other words, the categorization itself is flawed.
-
-%- Try to apply their model to a purely mixed state of spin 1/2
-%- Show that you can't have a single epistemic distribution
-%-> Failure of model
 
@@
 
-Let us now step back a moment. Why does assuming $p(\lambda|P)$ is well defined lead to an epistemic structure incompatible with the quantum one?
+Let us now ask a crucial question: Why does assuming that $p(\lambda|P)$ is well defined lead to an epistemic structure incompatible with QM?
 
-It is widely understood that one feature of quantum mechanics is contextuality: a measurement cannot be understood as revealing pre-existing values.
+It is widely understood that one feature of quantum theory is contextuality: a measurement cannot be conceived as revealing pre-existing values.
 We may be tempted to think that we have one context before and one after the measurement.
@@
 In the same way that we cannot understand quantum measurements as selecting from a single joint probability distribution for all properties, we cannot understand quantum preparations as constructing a single joint probability distribution over all properties.
-To fully understand how this works, let us review how probability spaces work in both classical and quantum mechanics.
+To fully understand how this works, let us review how probability spaces behave in both classical and quantum mechanics.
 
-A ``classical'' probability space is made of three objects: a sample space $\Omega$ which represents all possible cases (e.g.
-all points in phase-space for a classical system); a $\sigma$-algebra $\Sigma_\Omega$ over $\Omega$ which represents all statements of interest (e.g.
-``the position is between 2 and 3 meters''); a measure $\mu : \Sigma_\Omega \to \mathbb{R}$ that associates each statement with a probability.
+A ``classical'' probability space is made of three objects: a sample space $\Omega$ representing all possible cases (e.g.\ all points in phase-space for a classical system); a $\sigma$-algebra $\Sigma_\Omega$ over $\Omega$  representing all statements of interest (e.g.\ ``the position is between 2 and 3 meters''); a measure $\mu : \Sigma_\Omega \to \mathbb{R}$ associating each statement with a probability.
 Note how we have a single probability distribution over the whole space.
 
-In quantum mechanics we start with a similar structure: a set of states $\mathcal{H}$ which represents all possible cases (i.e.
-the Hilbert space) and a $\sigma$-algebra $\Sigma_{\mathcal{H}}$ that represents all statements of interest.
+In QM we start with a similar structure: a set of states $\mathcal{H}$ representing all possible cases (i.e.\ the Hilbert space) and a $\sigma$-algebra $\Sigma_{\mathcal{H}}$ representing all statements of interest.
 This is where the similarity ends.
-To specify a mixed state, a ``quantum distribution'', we use a density matrix $\rho : \mathcal{H} \to \mathcal{H}$.
+To specify a mixed state---a ``quantum distribution''---we use a density matrix $\rho : \mathcal{H} \to \mathcal{H}$.
 To get a probability distribution, we need to specify an observable $O$.
@@
 Mathematically, what we call context is a $\sigma$-algebra: a set of propositions upon which we assign probabilities.
-This tells us why we cannot simply use classical rules in quantum mechanics.
+This fact explains why we cannot simply use classical rules in QM.
 Classical mechanics has only one context, so we can always mix and match statements however we please.
-In quantum mechanics, propositions about different observables live in different $\sigma$-algebras, in different probability spaces, and therefore cannot be combined.
+In QM, propositions about different observables live in different $\sigma$-algebras---in different probability spaces---and therefore cannot be combined.
 
@@
 And this seems to work in some cases.
-Given any density matrix $\rho$, there exists a (non-unique) privileged context: this is the one in which we can express $\rho = \sum p_i |\psi_i > < \psi_i|$ as the mix of orthogonal pure states.
+Given any density matrix $\rho$, there exists a (non-unique) privileged context: this is the one in which we can express $\rho = \sum p_i |\psi_i \rangle \langle \psi_i|$ as the mix of orthogonal pure states.
 If $\rho$ is itself a pure state, any basis that includes it will be a privileged context.
@@
 
-Suppose we start with two density matrices $\rho_1$ and $\rho_2$ with different eigenstates.
-We then create the mixture $\rho = p_1 \rho_1 + p_2 \rho_2$.
+Suppose to consider two density matrices $\rho_1$ and $\rho_2$ with different eigenstates and then to create the mixture $\rho = p_1 \rho_1 + p_2 \rho_2$.
 The context of the mixture will not be the same as the original contexts.
@@
 
-To sum up, assuming $p(\lambda|P)$ exists means assuming there is a single probability space, a single $\sigma$-algebra, a single set of propositions, to which all probabilities are assigned.
+To sum up, assuming that $p(\lambda|P)$ exists means assuming that there is a single probability space, a single $\sigma$-algebra, a single set of propositions to which all probabilities are assigned.
 This is exactly assuming non-contextuality.
-While this can be made to work for pure state to pure state transitions during measurement, mixing preparations leads to an epistemic structure that is fundamentally non-contextual, leading to inconsistencies with quantum information.
+While this can be made to work for pure state-to-pure state transitions during measurements, mixing preparations leads to an epistemic structure that is fundamentally non-contextual, leading to inconsistencies with quantum information.
 To allow for contextuality we should assume that for each observable $O$ we have a probability $p_O(\lambda_O|P)$ of having prepared one of the eigenstates, such that $\sum_{\lambda_O} p_O(\lambda_O|P) = 1$.
@@
 
-% (mezza pagina)
-%- Non-contestuality of HS
-%- PBR and other results
-
 \section{Conclusion}
 
-%\section*{Supplementary material}
-%
-%TODO: do we need to bump anything here?
+%Number of words for the conclusion: 159
 
-\bibliographystyle{apalike}
+%\bibliographystyle{ieeetr}
 \bibliography{PhDthesis}

### 2022-01-17 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: another pass, added caption and tentative conclusions

@@
 \begin{abstract}
-The ontological model framework introduced by Harrigan and Spekkens has been widely used to categorize quantum interpretations and as a basis for influential results such as the PBR theorem.
-In this paper we show that the model has a fundamental flaw: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics.
-Namely, the map between the epistemic states of the model and quantum density matrices is not even an order homomorphism in terms of the information entropy, consequently the epistemic content of mixed states is not mapped in a meaningful way.
-Moreover, by recasting the definitions of the model in terms of density matrices, we find that only $\psi$-epistemic interpretations are allowed by quantum mechanics, a result that is the opposite of the PBR theorem.
-We show that the cause of this mismatch is the effect of non-contextuality over mixed preparations that does not allow for a single decomposition of mixed states over pure states.
+The Pusey, Barrett and Rudolph (PBR) theorem concludes that quantum states cannot be taken to represent mere information about the system.
+This result is based on the ontological model framework proposed by Harrigan and Spekkens (HS).
+In this paper we show that the HS framework has a fundamental problem: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics.
+Namely, the map between the epistemic states of the model and quantum density matrices preserves neither the value nor ordering of the information entropy.
+Consequently the epistemic content of mixed states is not mapped in a meaningful way.
+The problem stems from the assumption that an epistemic state is characterized by a \emph{single} probability measure, which is essentially an assumption of non-contextuality.
+Given this fundamental issue, every result that uses the HS framework, including the PBR theorem, should be carefully reexamined.
 \end{abstract}
@@
 
-Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b}, \cite{Renner:2012}, \cite{Colbeck:2017}, \cite{Hardy:2013}, \cite{Maroney:2014}, \cite{Patra:2013}, \cite{Mansfield:2016}, \cite{Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014}, \cite{Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information.
-On the other, scholars recently shown that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017}, \cite{Rizzi:2018}, \cite{Oldofredi:2021}, \cite{DeBrota:2019}).
+Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information.
+On the other, scholars recently showed that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2021, DeBrota:2019}).
 
-In this essay we are not going to rebut the theorem itself.
-We instead aim to carefully re-examine one of its premises, i.e.\ Harrigan and Spekkens (HS) classification between $\psi-$ontic and $\psi-$epistemic ontological models  \cite{Harrigan:2010} on top of which the PBR result is derived.
-We believe that a careful examination of this assumption forces us to re-evaluate the significance of the PBR conclusion.
+In this paper we are not going to rebut the theorem itself.
+We instead aim to carefully reexamine one of its premises, i.e.\ Harrigan and Spekkens' (HS) classification between $\psi-$ontic and $\psi-$epistemic ontological models  \cite{Harrigan:2010} on top of which the PBR result is derived.
+We believe that a careful examination of this assumption forces us to reevaluate the significance of the PBR conclusion.
 
-HS proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in that case the model is called $\psi-$ontic, or to some observer information, making it $\psi-$epistemic.
-While the original aim of this classification was to clarify Einstein's view of quantum theory, HS framework has been widely employed in the literature not only to categorize different interpretations, but also to argue what types of interpretations are admissible (\cite{Leifer:2013}, \cite{Leifer:2017}, \cite{Branciard:2014} \cite{Hermens:2021}, \cite{Wood:2015}, \cite{Ringbauer:2015}, \cite{Mazurek:2016}, \cite{Bartlett:2012}; cf.\ \cite{Oldofredi:2020b}, \cite{Lewis:2012}, \cite{Ladyman:2021} for critical discussions).
-Given the influence that HS framework and the PBR theorem have in the quantum foundations, it is worthwhile to look again at the former with more scrutiny in order to understand whether it is a sound basis to draw conclusions on the nature of the quantum state.
+HS proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi-$ontic, or to some observer information, making it $\psi-$epistemic.
+While the original aim of this classification was to clarify Einstein's view of quantum theory, the HS framework has been widely employed in the literature not only to categorize different interpretations, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Lewis:2012, Ladyman:2021} for critical discussions).
+Given the influence that the HS framework and the PBR theorem have in quantum foundations, it is worthwhile to look again at the former with more scrutiny in order to understand whether it is a sound basis from which to draw conclusions on the nature of the quantum state.
 
@@
 \begin{enumerate}
-	\item the information entropy as calculated on epistemic states does not match the one provided by quantum mechanics, meaning that the epistemic structures provided by QM and HS are incompatible;
-	\item the density matrices associated with pure states overlap for non-orthogonal states (i.e.
-$\rho_\psi \rho_\phi \neq 0$), meaning that $\psi$-ontic models are ruled out, contrary to the PBR claim.
+	\item the information entropy as calculated on HS epistemic states does not match the one provided by quantum mechanics, therefore the epistemic structures provided by QM and HS are incompatible;
+	\item in QM the density matrices associated with pure states overlap for non-orthogonal states (i.e. $\rho_\psi \rho_\phi \neq 0$), which rules out $\psi$-ontic models
 \end{enumerate}
-Hence, our conclusion is that HS distinction is fundamentally flawed.
-The main problem lies in assuming that the epistemic state identifies a single probability distribution over the whole set of ontic states.
+Combining these results with PBR, our conclusion is that the HS classification itself is fundamentally problematic.
+The main issue lies in assuming that the epistemic state identifies a \emph{single} probability distribution over the whole set of ontic states.
 We show that this, in essence, is an assumption of non-contextuality, which is also inherited by the PBR theorem, casting doubts on its validity.
@@
 We briefly review the main features of the classification provided by HS starting with the usual operational setting for QM.
-We have a preparation protocol $P$, which will be associated with a density operator $\rho$ on the relevant Hilbert space, and a measurement protocol $M$, represented by a POVM $\{ E_k\}$ where each $k$ represents a possible measurement outcome.
+We have a preparation protocol $P$, associated with a density operator $\rho$ on the relevant Hilbert space, and a measurement protocol $M$, represented by a POVM $\{ E_k\}$ where each $k$ represents a possible measurement outcome.
 The probability of obtaining a particular $k$ given a particular preparation $P$ and measurement $M$ is given by the generalized Born rule
@@
 Let us consider more closely the space of epistemic states given by $p(\lambda | P)$ and equation \ref{epistemic_mixing}.
-To make the discussion more concrete, we assume to have a $\psi$-complete model of a single qubit representing the direction of spin.
+To make the discussion more concrete, we assume we have a $\psi$-complete model of a single qubit representing the direction of spin.
 
@@
 \includegraphics[scale=.4]{fig2}
-\caption{\footnotesize{}}\label{fig_map}
+\caption{\footnotesize{The function $\iota$ maps the epistemic states $E_{HS}$ of a $\psi$-ontic model of a single qubit to the respective density matrices $E_Q$.
+Note that neither the value nor the ordering of the information entropy is preserved.}}\label{fig_map}
 \end{figure}
@@
 In the same way that we cannot understand quantum measurements as selecting from a single joint probability distribution for all properties, we cannot understand quantum preparations as constructing a single joint probability distribution over all properties.
-To fully understand how this works, let us review how probability spaces behave in both classical and quantum mechanics.
+To fully understand how this works, let us review how probability spaces are defined in both classical and quantum mechanics.
 
@@
 
-Suppose to consider two density matrices $\rho_1$ and $\rho_2$ with different eigenstates and then to create the mixture $\rho = p_1 \rho_1 + p_2 \rho_2$.
+Consider two density matrices $\rho_1$ and $\rho_2$ with different eigenstates and their mixture $\rho = p_1 \rho_1 + p_2 \rho_2$.
 The context of the mixture will not be the same as the original contexts.
@@
 
-To sum up, assuming that $p(\lambda|P)$ exists means assuming that there is a single probability space, a single $\sigma$-algebra, a single set of propositions to which all probabilities are assigned.
+To sum up, assuming $p(\lambda|P)$ exists means assuming there is a single probability space, a single $\sigma$-algebra, a single set of propositions to which all probabilities are assigned.
 This is exactly assuming non-contextuality.
@@
 
+While work in quantum foundations typically focuses on measurements and their effect on pure states, equal importance must be given to preparations and their effect on mixed states.
+The different behavior of entropy in classical and quantum mechanics has profound implications not just for information theory, but also for statistical mechanics and its measurable predictions.
+A full solution to the ``measurement problem'' will therefore have to address the dual ``preparation problem.''
+
 %Number of words for the conclusion: 159

### 2022-01-20 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: final from Andrea

@@
-\documentclass[twocolumn,prl,floatfix,superscriptaddress]{revtex4-2}
+\documentclass[twocolumn,prl,floatfix,superscriptaddress]{revtex4-1}
 \usepackage{graphicx, subfigure, amsfonts, amssymb, amsmath, array, gensymb, fontenc, color, lipsum}
@@
 \section{Introduction}
-
-%---e.g.\ standard Quantum Mechanics (QM) is classified as $\psi$-ontic and $\psi-$complete, hidden variable theories such as Bohm causal approach \cite{Bohm:1952aa} and Nelson mechanics \cite{Nelson:1966aa} are both considered $\psi$-ontic and $\psi$-incomplete, whereas Relational QM \cite{Rovelli:1996} and Quantum Bayesianism \cite{Fuchs:2002} are $\psi-$epistemic.
-%Since its publication HS paper has been subjected to several analyses and extensions as well as critical discussions showing some limitations of this approach (\cite{Leifer:2014, Leifer:2014b}, \cite{Leifer:2013}, \cite{Hermens:2021}, \cite{Wood:2015}, \cite{Ringbauer:2015}, \cite{Mazurek:2016}, \cite{Bartlett:2012}, \cite{Oldofredi:2020b}, \cite{Lewis:2012}, \cite{Ladyman:2021}).
-However, still today HS distinction is employed to argue what types of interpretations are possible.
-Referring to this, the PBR theorem \cite{PBR:2012}---a remarkable result derived on top of HS framework---rules out $\psi$-epistemic models.
-More precisely, PBR argued that in every model reproducing the statistics and predictions of QM the quantum state $\psi$ must represent real physical properties of the system under consideration---i.e.\ models must be $\psi-$ontic.
-Consequently, quantum theories cannot be $\psi-$epistemic.
-Given the importance of this claim and the resonance that such a theorem had (\cite{Renner:2012}, \cite{Hardy:2013}, \cite{Maroney:2014}, \cite{Patra:2013}, \cite{Mansfield:2016}, \cite{Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014}, \cite{Ben:2017}, \cite{Rizzi:2018}, \cite{Oldofredi:2021}, \cite{DeBrota:2019}), it is worthwhile to look at HS framework again with more scrutiny in order to understand whether it is a technically sound starting point to draw conclusions on the nature of the quantum state.
 
@@
 Temperature is a property of the preparation and therefore an \emph{epistemic property}.
-
-%While the original aim of this classification was to clarify Einstein's view of quantum theory, it has been used in the literature to classify different interpretations (e.g.
-hidden variable theories such as Bohm pilot-wave theory \cite{Bohm:1952aa} and Nelson mechanics \cite{Nelson:1966aa} are both considered $\psi$-ontic and $\psi$-incomplete) and also to argue what types of interpretations are possible.
-Most notably, the PBR theorem \cite{PBR:2012} rules out any $\psi$-epistemic model.
-More precisely, PBR argued that in every model reproducing the statistics and predictions of Quantum Mechanics (QM) the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi-$ontic.
-Consequently, quantum theories cannot be $\psi-$epistemic.(TODO beef up this paragraph with citations).
-It is therefore worthwhile to look at the ontological model itself with more scrutiny.
 
@@
 To allow for contextuality we should assume that for each observable $O$ we have a probability $p_O(\lambda_O|P)$ of having prepared one of the eigenstates, such that $\sum_{\lambda_O} p_O(\lambda_O|P) = 1$.
-While we have a single state space $\Lambda$, we do not have a single probability space over the whole space.
+Although we have a single state space $\Lambda$, we do not have a single probability space over the whole space.
 The entropy of the preparation would be the lowest entropy across all contexts.
@@
 
-While work in quantum foundations typically focuses on measurements and their effect on pure states, equal importance must be given to preparations and their effect on mixed states.
+In this essay we showed not only that HS framework does not correctly describe the epistemic structure of QM, but also that it implicitly assumes non-contextuality.
+Hence, it cannot be employed to draw conclusions on the nature of quantum states.
+Consequently, any other result relying on HS categorization inherits the same issues affecting it.
+In particular, it follows that the PBR theorem is derived from a technically ill-defined basis; thus, it is demonstrated neither that quantum states must necessarily be HS-$\psi-$ontic, nor that epistemic interpretations are contradicted by quantum theory.
+Concluding, we believe that while work in quantum foundations typically focuses on measurements and their effect on pure states, equal importance must be given to preparations and their effect on mixed states.
 The different behavior of entropy in classical and quantum mechanics has profound implications not just for information theory, but also for statistical mechanics and its measurable predictions.
-A full solution to the ``measurement problem'' will therefore have to address the dual ``preparation problem.''
 
-%Number of words for the conclusion: 159
+%A full solution to the ``measurement problem'' will therefore have to address the dual ``preparation problem.''
+
+%Number of words for the conclusion: 164
+
+\section{Acknowledgements}
+
+Andrea Oldofredi is grateful to the Funda\c{c}$\tilde{\mathrm{a}}$o para a Ci\^encia e a Tecnologia (FCT) for financial support (Grant no. 2020.02858).
 

### 2022-01-22 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: another pass, mainly minor typos

@@
-\documentclass[twocolumn,prl,floatfix,superscriptaddress]{revtex4-1}
+\documentclass[twocolumn,prl,floatfix,superscriptaddress]{revtex4-2}
 \usepackage{graphicx, subfigure, amsfonts, amssymb, amsmath, array, gensymb, fontenc, color, lipsum}
@@
 \affiliation{Centre of Philosophy (CFUL), University of Lisbon, Portugal}
-\author{Christine Aidala}
+\author{Christine A. Aidala}
 \affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
@@
 In 2012 Pusey, Barrett and Rudolph published a formal result in \emph{Nature Physics}, widely known as the PBR theorem, showing that ``if the quantum state merely represents information about the real physical state of a system, then experimental predictions are obtained that contradict those of quantum theory'' (\cite{PBR:2012}, p.\ 475).
-Alternatively stated, PBR argued that in every model reproducing the statistics and predictions of Quantum Mechanics (QM) the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi-$ontic.
-Consequently, quantum theories cannot be $\psi-$epistemic.
+Alternatively stated, PBR argued that in every model reproducing the statistics and predictions of Quantum Mechanics (QM) the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi$-ontic.
+Consequently, quantum theories cannot be $\psi$-epistemic.
 
@@
 In this paper we are not going to rebut the theorem itself.
-We instead aim to carefully reexamine one of its premises, i.e.\ Harrigan and Spekkens' (HS) classification between $\psi-$ontic and $\psi-$epistemic ontological models  \cite{Harrigan:2010} on top of which the PBR result is derived.
+We instead aim to carefully reexamine one of its premises, i.e.\ Harrigan and Spekkens' (HS) classification between $\psi$-ontic and $\psi$-epistemic ontological models  \cite{Harrigan:2010} on top of which the PBR result is derived.
 We believe that a careful examination of this assumption forces us to reevaluate the significance of the PBR conclusion.
 
-HS proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi-$ontic, or to some observer information, making it $\psi-$epistemic.
+HS proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic.
 While the original aim of this classification was to clarify Einstein's view of quantum theory, the HS framework has been widely employed in the literature not only to categorize different interpretations, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Lewis:2012, Ladyman:2021} for critical discussions).
@@
 \includegraphics[scale=.7]{ontic}
-\caption{\footnotesize{Harrigan and Spekkens distinction between $\psi-$ontic (a) and $\psi-$epistemic (b) ontological models.}}\label{overlap}
+\caption{\footnotesize{Harrigan and Spekkens' distinction between $\psi$-ontic (a) and $\psi$-epistemic (b) ontological models.}}\label{overlap}
 \end{figure}
@@
 This is a direct consequence of defining the epistemic states as a distribution $p(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability to each ontological state $\lambda$ \footnote{We will follow the convention of not distinguishing between probability and probability density.}.
-In QM, mixed states $E_{Q}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
+In QM, mixed states $E_{QM}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
 Looking at the difference in dimension, the discrepancy is not minor.
@@
 The difference is that, in the quantum case, we have an ``aliasing'': different mixtures will give the same mixed states.
-Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{Q}$ that allows us to go from HS-epistemic states to density matrices, the Q-epistemic states.
+Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{QM}$ that allows us to go from HS-epistemic states to density matrices, the QM-epistemic states.
 This map is surjective but not injective.
-Therefore $E_{Q}$ can be seen as a set of equivalence classes of $E_{HS}$.
-The key question, then, is the following: is the map $\iota$ sufficiently well behaved that we can understand the epistemic content of $E_{Q}$ based on the epistemic content of $E_{HS}$?
+Therefore $E_{QM}$ can be seen as a set of equivalence classes of $E_{HS}$.
+The key question, then, is the following: is the map $\iota$ sufficiently well behaved that we can understand the epistemic content of $E_{QM}$ based on the epistemic content of $E_{HS}$?
 Or is it so ill behaved that there is no connection between the two?
 
-Ideally, one would like to say that $E_{HS}$ represents all possible epistemic states that include the knowledge about preparation, while $E_{Q}$ represents only those distinguishable through measurement.
+Ideally, one would like to say that $E_{HS}$ represents all possible epistemic states that include the knowledge about preparation, while $E_{QM}$ represents only those distinguishable through measurement.
 For example, one may want to say that the axis of preparation is, in principle, an ontologically distinct property, but this is not experimentally distinguishable.
-Therefore, distinct HS-epistemic states for each direction are projected by $\iota$ onto the same Q-epistemic state of maximum uncertainty.
+Therefore, distinct HS-epistemic states for each direction are projected by $\iota$ onto the same QM-epistemic state of maximum uncertainty.
 This sounds very plausible and appealing, but it turns out this is not the case.
 To show this fact we will concentrate on a single aspect of the map: how information entropy transforms under $\iota$.
-While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is sufficient to show that the epistemic structures of $E_{HS}$ and $E_{Q}$ are different.
+While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is sufficient to show that the epistemic structures of $E_{HS}$ and $E_{QM}$ are different.
 
@@
 Since $\iota(A)$ will be the maximally mixed state, this also corresponds to entropy 1.
-So far so good: $\iota$ maps entropy 1 to entropy 1!
+So far so good: $\iota$ maps entropy 1 to entropy 1.
 However, note that \textbf{A} is a measure zero set of \textbf{B}, which is a measure zero set of \textbf{C}.
@@
 Note that, in this case, $\iota$ picks the minimum entropy between the equivalent epistemic states.
-This is inconsistent with the idea that $E_{Q}$ represents states where knowledge is lost: that would pick the maximum possible entropy.
+This is inconsistent with the idea that $E_{QM}$ represents states where knowledge is lost: that would pick the maximum possible entropy.
 
@@
 \includegraphics[scale=.4]{fig2}
-\caption{\footnotesize{The function $\iota$ maps the epistemic states $E_{HS}$ of a $\psi$-ontic model of a single qubit to the respective density matrices $E_Q$.
+\caption{\footnotesize{The function $\iota$ maps the epistemic states $E_{HS}$ of a $\psi$-ontic model of a single qubit to the respective density matrices $E_{QM}$.
 Note that neither the value nor the ordering of the information entropy is preserved.}}\label{fig_map}
@@
 However, $\iota(D)$ is not the maximally mixed state and therefore the entropy is less than 1.
-This means that epistemic states that are isoentropic in $E_{HS}$ are not isoentropic in $E_{Q}$.
-If we think as $E_{HS}$ and $E_{Q}$ as partially ordered by information entropy, then $\iota$ is not an order preserving map.
-Therefore, the epistemic structures represented by $E_{HS}$ and $E_{Q}$ are not isomorphic.
+This means that epistemic states that are isoentropic in $E_{HS}$ are not isoentropic in $E_{QM}$.
+If we think of $E_{HS}$ and $E_{QM}$ as partially ordered by information entropy, then $\iota$ is not an order preserving map.
+Therefore, the epistemic structures represented by $E_{HS}$ and $E_{QM}$ are not isomorphic.
 
@@
 
-In this essay we showed not only that HS framework does not correctly describe the epistemic structure of QM, but also that it implicitly assumes non-contextuality.
+We showed not only that the HS framework does not correctly describe the epistemic structure of QM, but also that it implicitly assumes non-contextuality.
 Hence, it cannot be employed to draw conclusions on the nature of quantum states.
-Consequently, any other result relying on HS categorization inherits the same issues affecting it.
-In particular, it follows that the PBR theorem is derived from a technically ill-defined basis; thus, it is demonstrated neither that quantum states must necessarily be HS-$\psi-$ontic, nor that epistemic interpretations are contradicted by quantum theory.
+Consequently, any other result relying on the HS categorization inherits the same issues.
+In particular, it follows that the PBR theorem is derived from a technically ill-defined basis; thus, it is demonstrated neither that quantum states must necessarily be HS-$\psi$-ontic, nor that epistemic interpretations are contradicted by quantum theory.
 Concluding, we believe that while work in quantum foundations typically focuses on measurements and their effect on pure states, equal importance must be given to preparations and their effect on mixed states.
@@
 
-Andrea Oldofredi is grateful to the Funda\c{c}$\tilde{\mathrm{a}}$o para a Ci\^encia e a Tecnologia (FCT) for financial support (Grant no. 2020.02858).
+Andrea Oldofredi is grateful to the Funda\c{c}$\tilde{\mathrm{a}}$o para a Ci\^encia e a Tecnologia (FCT) for financial support (Grant no.
+2020.02858).
+This work is in connection to Assumptions of Physics, a larger project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).
 
 %\bibliographystyle{ieeetr}
-\bibliography{PhDthesis}
+\bibliography{HarriganSpekkens}
 

### 2022-01-27 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
HarriganSpekkens: to submit to archives

@@
 \author{Andrea Oldofredi}
-\affiliation{Centre of Philosophy (CFUL), University of Lisbon, Portugal}
+\affiliation{Centre of Philosophy, University of Lisbon, Portugal}
 \author{Christine A. Aidala}
@@
 \begin{abstract}
-The Pusey, Barrett and Rudolph (PBR) theorem concludes that quantum states cannot be taken to represent mere information about the system.
+The Pusey, Barrett and Rudolph (PBR) theorem claims that quantum states cannot be taken to represent mere information about the system.
 This result is based on the ontological model framework proposed by Harrigan and Spekkens (HS).
@@
 \begin{enumerate}
-	\item the information entropy as calculated on HS epistemic states does not match the one provided by quantum mechanics, therefore the epistemic structures provided by QM and HS are incompatible;
-	\item in QM the density matrices associated with pure states overlap for non-orthogonal states (i.e. $\rho_\psi \rho_\phi \neq 0$), which rules out $\psi$-ontic models
+	\item the information entropy as calculated on HS epistemic states does not match the one provided by quantum mechanics, therefore the epistemic structures provided by QM and HS are not isomorphic;
+	\item if we rephrase the overlapping condition directly in terms of density matrices, we find that non-orthogonal pure states ``overlap'', which rules out $\psi$-ontic models.
 \end{enumerate}
@@
 \end{equation}
-As the epistemic states for mixed preparations are simply linear combinations of epistemic states for pure preparations, we can concentrate on the latter.
+As the epistemic states for mixed preparations are simply linear combinations of epistemic states for pure preparations, the HS model concentrates on the latter.
 
@@
 Note that such epistemic distributions do not overlap.
-Therefore, according to HS categorization, these correspond to two ontologically distinct states, making the preparation axis an ontological property of the system, regardless of the orientation.
+Therefore, according to HS categorization, these correspond to two ontologically distinct states, making the preparation axis an ontological property \emph{of the system}, regardless of the orientation.
 However, the respective density matrices would not just overlap: they would be the same maximally mixed state.
@@
 \begin{align}\label{conclusion}
-	\parbox{2.8in}{Non-orthogonal pure quantum states overlap and are therefore epistemic.}
+	\parbox{2.8in}{Non-orthogonal pure quantum states overlap and are therefore HS-epistemic.}
 \end{align}
@@
 Andrea Oldofredi is grateful to the Funda\c{c}$\tilde{\mathrm{a}}$o para a Ci\^encia e a Tecnologia (FCT) for financial support (Grant no.
-2020.02858).
+2020.02858.CEECIND).
 This work is in connection to Assumptions of Physics, a larger project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived  (\url{https://assumptionsofphysics.org}).

### 2022-01-28 TeachingMaterial/ActionPrinciple/ActionPrinciple.tex
ActionPrinciple: first draft

@@
 \begin{abstract}
-TBD
+We give a geometrical interpretation for the action principle in Lagrangian classical particle mechanics that also has a direct physical meaning.
+The idea is that the variation of the action along a path is the flow of the displacement field through the area between the path and the variation.
+Since actual trajectories are parallel to the displacement field, the flow is zero along them, making the action stationary.
+We first give the interpretation for a single degree of freedom using simple differential calculus so that it is accessible to the widest possible audience.
+We then generalize to multiple degrees of freedom using symplectic geometry.
+This treatment makes the geometry of both Hamiltonian and Lagrangian mechanics more readily evident, and gives reasons to privilege the Hamiltonian picture over the Lagrangian one.
 \end{abstract}
@@
 
-The action principle is of paramount importance in physics, yet it is somewhat mysterious.
-It is, in essence, a geometrical relationship yet it is not clear where and why that relationship originates.
-The fact that the Lagrangian for a system is not unique, makes it even more difficult to understand what is the physical content it provides.
+While the principle of stationary action is regarded by many as one of the most important tools in physics, its physical meaning is not completely clear.
+First of all, the typical characterization of the Lagrangian as the difference between kinetic and potential energy fails even for simple systems, like a charged particle under a magnetic field.
+Moreover, the Lagrangian for a system is not uniquely defined, making the actual value of the action for a path not directly physically significant.
+We are left to wonder: what exactly is the action and why is it stationary for actual trajectories?
 
-In this paper we give a full characterization that is both geometrical and physical.
-We will see that the only physical assumption needed to recover both Hamiltonian mechanics and the action principle is that of deterministic and reversible evolution.
-We will also see how the action principle arises as a feature of divergence-free fields and closed two-forms.
-In a nutshell, we will see that the action principle arises as a general mathematical property of vector calculus and differential geometry, its physical significance is of less importance than the assumption of determinism and reversibility it requires.
+As part of our larger project Assumptions of Physics, we developed an approach, called Reverse Physics, which examines current theories to find a set of starting physical assumptions that are sufficient to rederive them.
+This allowed us to recast Hamiltonian mechanics as the dynamics of deterministic and reversible evolution, which coincidentally gives us the proper setting to characterize both the physics and the geometry underlying the principle of stationary action.
+We will see that the principle arises as a feature of divergence-free fields and closed two-forms, as a general mathematical property of vector calculus and differential geometry.
 
 We will first address the case of a single degree of freedom using standard vector calculus.
-This provides already all the geometrical and physical insights in a language that is accessible to most physicist.
-Then we proceed to the case of multiple independent degrees of freedom using differential geometry.
-These new tools, though unfamiliar to many physicists and engineers, are unfortunately needed.
-Yet, we will give it a treatment and a notation that hopefully will make them more intuitive.
+This provides already all the geometrical and physical insights in a language that is accessible to all physicists and engineers.
+Then we proceed to the case of multiple independent degrees of freedom, which requires tools from differential geometry.
+To make the treatment of the latter case more accessible, we will use a notation that is similar to the one used by physicists in general relativity.
 
 \section{One degree of freedom}
-
-We are going to present the simple case of one degree of freedom using the standard tools provided by vector calculus.
-It is very fortunate that this can be done at all, since it allows the majority of people with a physics or engineering background to grasp all the major points.
 
@@
 That is, the space charted by position $q$, momentum $p$ and time $t$.
-It is precisely because we have only three variable that make the standard tools of vector calculus appropriate.
+It is precisely because we have exactly three variables that the standard tools of vector calculus can be used.
 
-FIGURE: a few trajectories in extended phase space
+\begin{figure}
+	\includegraphics[width = 0.45\textwidth]{ExtendedPhaseSpace}
+	\caption{\footnotesize{Evolutions in the extended phase space and the divergence-free displacement field.}}\label{extended_phase_space}
+\end{figure}
 
-Once we have recast the equation of motion on the extended phase space, we can see how the action principle comes from a straightforward application of vector calculus in the case of divergence-free fields.
+Once we have recast the equations of motion on the extended phase space, we can see how the action principle comes from a straightforward application of vector calculus in the case of divergence-free fields.
 
@@
 
-Under this assumption, given a state at a moment in time, there is a well-defined change in all variable which is described by the displacement vector field
+Under this assumption, given a state at a moment in time, there is a well-defined change in all variables which is described by the displacement vector field
 \begin{equation}
@@
 to properly represent a deterministic and reversible evolution.
-
-FIGURE: region in extended phase space with trajectories flowing through
 
@@
 Given that $\vec{\theta}$ is a vector potential, we have the usual gauge arbitrariness since $\nabla \times(\vec{\theta} + \nabla f) = \nabla \times \vec{\theta}$ and therefore the displacement field $\vec{S}$ remains unchanged.
-By choosing $f$ appropriately, we can set
+By choosing $f(q,p,t)$ appropriately, we can set
 \begin{equation}
@@
 \end{align*}
-Integrating, we have $\theta_q = p + g(q,t)$ where $g(q,t)$ is an arbitrary function, which we can set to zero. Therefore we have:
+Integrating, we have $\theta_q = p + g(q,t)$ where $g(q,t)$ is an arbitrary function.
+This function can be set to zero without loss of generality as it can be removed with a gauge transformation where $f(q,t)$ does not depend on $p$, and therefore $\theta_p$ will remain unchanged.
+Therefore we have:
 \begin{equation}
@@
 In the first case we are taking a total derivative along the evolution, and therefore the momentum can change.
-In the second case we are taking a partial derivative which is taken at constant $q$ and $p$ by definition, and therefore $\frac{\partial p}{\partial t}=0$.}
+In the second case we are taking a partial derivative at constant $q$ and $p$ by definition, and therefore $\frac{\partial p}{\partial t}=0$.}
 \begin{align*}
@@
 
-Let us now review the principle of least action.
+Let us now review the principle of stationary action.
 We chart all states at all times by using $\{q, \frac{dq}{dt}, t\}$.
-We have a Langrangian $L(q, \frac{dq}{dt}, t)$ and, given a path $\gamma = \{q(t), \frac{dq}{dt}(t), t\}$ with $t_1 \leq t \leq t_2$, we define the action as $\mathcal{A}[\gamma] = \int_{t_1}^{t_2} L(q, \frac{dq}{dt}, t) dt$.
-An actual evolution of the system is such that the action is stationary:
+We have a Langrangian $L(q, \frac{dq}{dt}, t)$ and, given a path $\gamma = \{q(t), \frac{dq}{dt}(t), t\}$ with $t_1 \leq t \leq t_2$, we define the action as
+\begin{equation}
+	\mathcal{A}[\gamma] = \int_{t_1}^{t_2} L(q, \frac{dq}{dt}, t) dt.
+\end{equation}  An actual evolution of the system is such that the action is stationary:
 \begin{equation}
@@
 \end{equation}
-The connection between the Lagrangian and Hamiltonian formulation is given by
+The connection between the Lagrangian and Hamiltonian formulation is given by the Legendre transformation
 \begin{equation}
@@
 Let us step back, forget about Lagrangians, and resume the discussion from before.
-Let $\gamma$ be an evolution (i.e.
-a solution of the equations of motion) with endpoints $A$ and $B$.
+Let $\gamma$ be a path in the extended phase space, not necessarily a solution of the equations of motion, with endpoints $A$ and $B$.
+That is, the displacement $d\vec{\gamma}$ along the curve will not be, in general, equal to $\vec{S}dt$.
 Consider the line integral of the vector potential $\vec{\theta}$ along $\gamma$.
@@
 \begin{aligned}
-	\int_A^B \vec{\theta} \cdot d\gamma &= \int^{t_1}_{t_0} \vec{\theta} \cdot \vec{S} dt \\
-	&= \int^{t_1}_{t_0} \left(\theta_q \frac{dq}{dt} + \theta_p \frac{dp}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
-	&= \int^{t_1}_{t_0} \left(p \frac{dq}{dt} + 0 \frac{dp}{dt} - H \frac{dt}{dt}\right) dt \\
-	&= \int^{t_1}_{t_0} \left(p \frac{dq}{dt} - H\right) dt
+	\int_A^B \vec{\theta} \cdot d\vec{\gamma} &= \int^{t_2}_{t_1} \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt} dt \\
+	&= \int^{t_2}_{t_1} \left(\theta_q \frac{dq}{dt} + \theta_p \frac{dp}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
+	&= \int^{t_2}_{t_1} \left(p \frac{dq}{dt} + 0 \frac{dp}{dt} - H \frac{dt}{dt}\right) dt \\
+	&= \int^{t_2}_{t_1} \left(p \frac{dq}{dt} - H\right) dt
 \end{aligned}
 \end{equation}
-In other words, the integral of the vector potential over an evolution coincides with the action, and the Lagrangian is simply
+In other words, the line integral of the vector potential over an evolution coincides with the action, and the Lagrangian is simply
 \begin{equation}
-	L = \vec{\theta} \cdot \vec{S}.
+	L = \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt}.
 \end{equation}
 
-FIGURE: trajectory with variation forming a closed loop. Another small area with trajectories flowing through
+\begin{figure}
+	\includegraphics[width = 0.45\textwidth]{ActionNonOptimized.png}
+	\includegraphics[width = 0.45\textwidth]{ActionOptimized.png}
+	\caption{\footnotesize{The variation of the action is the flow of the displacement field $\vec{S}$ through the surface $\Sigma$ that sits between the path $\gamma$ and its variation $\gamma'$.
+In the second picture we see that the flow is zero if the path is an actual evolution of the system, since the displacement field will be parallel to the path $\gamma$ and therefore the surface $\Sigma$.}}\label{action}
+\end{figure}
 
@@
 The original path $\gamma$ together with its variation $\gamma'$ form a contour $\partial \Sigma$ which encloses the surface $\Sigma$.
-We can therefore use Stoke's theorem to transform the line integral of $\vec{\theta}$ over $\partial \Sigma$ to a surface integral of the curl $\nabla \times \vec{\theta}$ over $\Sigma$.
+We can therefore use Stokes' theorem to transform the line integral of $\vec{\theta}$ over $\partial \Sigma$ to a surface integral of the curl $\nabla \times \vec{\theta}$ over $\Sigma$.
 That is:
 \begin{align*}
-	\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} = 
+	\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} 
 	&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma'} \\
-	&= \oint_{\partial \Sigma} \vec{\theta}  \cdot d\lambda \\
+	&= \oint_{\partial \Sigma} \vec{\theta}  \cdot d\vec{\gamma} \\
 	&= \iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
@@
 \end{align*}
-Since $\gamma$ is an evolution for the system, it is always tangent to $\vec{S}$.
-This means that the surface $\Sigma$ enclosed by $\gamma$ and its infinitesimal variation is also parallel to $\vec{S}$ and therefore $\vec{S} \cdot d\vec{\Sigma} = 0$.
-Which means
+Note that, precisely because $\vec{S}$ is divergence-free, it doesn't matter which particular surface $\Sigma$ we pick, as long as its contour $\partial \Sigma$ consists of $\gamma$ and $\gamma'$.
+
+The variation of the action, then, corresponds to the flow of $\vec{S}$ through $\Sigma$.
+It tells us how many actual evolutions go through the surface.
+Note that if $\gamma$ is an evolution of the system, it is always tangent to $\vec{S}$.
+This means that the surface $\Sigma$ enclosed by $\gamma$ and its infinitesimal variation is also tangent to $\vec{S}$ and therefore $\vec{S} \cdot d\vec{\Sigma} = 0$.
+Which means the actual evolutions are those paths $\gamma$ for which
 \begin{equation}
@@
 \end{equation}
-Which recover the action principle.
+This recovers the principle of stationary action.
 
@@
 Every evolution $\gamma$ is a field line of $\vec{S}$.
-Any infinitesimal surface $\Sigma$ enclosed by an evolution and its variation, then, will also be tangent to the displacement field $\vec{S}$ and therefore the surface integral of $\vec{S}$ over $\Sigma$ will be zero.
-Given that $\vec{S}$ is divergence-free, the integral over $\Sigma$ can be expressed as the line integral of the vector potential $\vec{\theta}$ over the contour of $\Sigma$.
+Any infinitesimal surface $\Sigma$ enclosed by an evolution and its variation, then, will also be tangent to the displacement field $\vec{S}$ and therefore the surface integral of $\vec{S}$ through $\Sigma$ will be zero.
+Given that $\vec{S}$ is divergence-free, the integral through $\Sigma$ can be expressed as the line integral of the vector potential $\vec{\theta}$ over the contour of $\Sigma$.
 Asking for a stationary action means asking for the flow of $\vec{S}$ to be zero across the surface given by the path and its variation, which happens only if the path is tangent to $\vec{S}$ and is therefore an evolution.
-In practice, it is a very roundabout way to ask that the path be tangent to the displacement vector.
+In practice, the principle of stationary action is a very roundabout way to ask that the path be always tangent to the displacement vector field.
 
@@
 The displacement field $\vec{S}$ and its properties are the only mathematical entities that are directly physically meaningful.
-The vector potential $\vec{\theta}$ is not uniquely defined by condition \ref{Detrev_assumptions}, though it is extremely useful as it allows us to map the space of all possible displacement fields of interested with a single ``scalar'' function.
+The vector potential $\vec{\theta}$ is not uniquely defined by condition \ref{Detrev_assumptions}, though it is extremely useful as it allows us to map the space of all possible displacement fields of interest with a single ``scalar'' function $H$, the time component of the vector potential.
 
@@
 In these cases, the Lagrangian approach cannot be applied because $\vec{\theta} \cdot \vec{S}$ cannot be expressed as a function of position and velocity.
-Yet, if we express $L(q,p,t)$ in terms of conjugate momentum, the modified version of the variational principle still works.
+Yet, if we express $L(q,p,t)$ in terms of conjugate momentum instead of velocity, the modified version of the variational principle still works.
 
@@
 \begin{enumerate}
-	\item the only Lagrangian systems in which the action principle yield a single solutions are the ones that admit a Hamiltonian formulation;
-	\item all Hamiltonian systems admit an action principle formulation, even though that do not admit a Lagrangian.
+	\item the only Lagrangian systems in which the action principle yields a single solution are the ones that admit a Hamiltonian formulation;
+	\item all Hamiltonian systems admit an action principle formulation, even if they do not admit a Lagrangian.
 \end{enumerate}
@@
 Given that we want to make the generalization understandable to a wide array of people with physical background, we are going to use enough of the standard terminology from differential geometry that our methods are recognizable to someone proficient in it, but we are going to use a notation that parallel Einstein's notation, so that it feels more familiar to a physicist.\footnote{The main differences are as follows.
-The vector basis will be noted as $e_a$ instead of $\frac{\partial}{\partial x_a}$.
+The vector basis will be noted as $e_a$ instead of $\frac{\partial}{\partial x^a}$.
 The covector basis will be noted as $e^a$ instead of $dx^a$.
@@
 \begin{aligned}
-	\omega(e^{q^i}, e^{p_j}) &= (-d\theta)_{q^i p_j} = -(\partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i}) = \delta^i_j \\
-	\omega_{q^i q^j} &= (-d\theta)_{q^i q^j} = -(\partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i}) = 0 \\
-	\omega(e^{p_i}, e^{p_j}) &= (-d\theta)_{p_i p_j} = -(\partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i}) = 0
+	\omega(e^{q^i}, e^{p_j}) &= (-\partial\theta)_{q^i p_j} = -(\partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i}) = \delta^i_j \\
+	\omega_{q^i q^j} &= (-\partial\theta)_{q^i q^j} = -(\partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i}) = 0 \\
+	\omega(e^{p_i}, e^{p_j}) &= (-\partial\theta)_{p_i p_j} = -(\partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i}) = 0
 \end{aligned}
@@
 Integrating, we have $\theta_{q^1} = p_1 + g(q^i, p_2, p_3, ..., t)$ where $g$ is an arbitrary function which we can set to zero.
+Like in the single degree of freedom, we can do this because $g$ does not depend on $p_1$, so it can be eliminated with a gauge transformation that does not change $\theta_{p_1}$.
 Therefore we have:
@@
 
-At this point, simply calculate $d\theta(S, \cdot ) $. The components of $d\theta$ are:
+At this point, we calculate $\partial\theta(S, \cdot ) $. The components of $\partial\theta$ are:
 \begin{equation*}
-	(d\theta)_{ab} = \begin{bmatrix}
+	(\partial\theta)_{ab} = \begin{bmatrix}
 		0 & \delta^j_i & - \partial_{q^i} H \\
@@
 \begin{align*}
-	d\theta(S, \cdot )  &= S^a (d\theta)_{ab} e^b = 0 \\
-	&= (S^{q^i}(d\theta)_{q^ib} + S^{p_i}(d\theta)_{p_ib} + S^{t}(d\theta)_{tb}) e^b \\
-	&= (S^{q^i}(d\theta)_{q^iq^j} + S^{p_i}(d\theta)_{p_iq^j} + S^{t}(d\theta)_{tq^j}) e^{q^j} + \\
-	& (S^{q^i}(d\theta)_{q^ip_j} +  S^{p_i}(d\theta)_{p_ip_j} + S^{t}(d\theta)_{tp_j}) e^{p_j} + \\
-	& (S^{q^i}(d\theta)_{q^it} + S^{p_i}(d\theta)_{p_it} + S^{t}(d\theta)_{tt}) e^t \\
+	\partial\theta(S, \cdot )  &= S^a (\partial\theta)_{ab} e^b = 0 \\
+	&= (S^{q^i}(\partial\theta)_{q^ib} + S^{p_i}(\partial\theta)_{p_ib} + S^{t}(\partial\theta)_{tb}) e^b \\
+	&= (S^{q^i}(\partial\theta)_{q^iq^j} + S^{p_i}(\partial\theta)_{p_iq^j} + S^{t}(\partial\theta)_{tq^j}) e^{q^j} + \\
+	& (S^{q^i}(\partial\theta)_{q^ip_j} +  S^{p_i}(\partial\theta)_{p_ip_j} + S^{t}(\partial\theta)_{tp_j}) e^{p_j} + \\
+	& (S^{q^i}(\partial\theta)_{q^it} + S^{p_i}(\partial\theta)_{p_it} + S^{t}(\partial\theta)_{tt}) e^t \\
 	&= (-S^{p_i}\delta^i_j + S^{t}\partial_{q^j} H ) e^{q^j} + \\

### 2022-01-30 TeachingMaterial/ActionPrinciple/ActionPrinciple.tex
ActionPrinciple: first draft ready

@@
 Let us step back, forget about Lagrangians, and resume the discussion from before.
-Let $\gamma$ be a path in the extended phase space, not necessarily a solution of the equations of motion, with endpoints $A$ and $B$.
-That is, the displacement $d\vec{\gamma}$ along the curve will not be, in general, equal to $\vec{S}dt$.
+As shown in fig.
+\ref{action}-a, let $\gamma$ be a path in the extended phase space with endpoints $A$ and $B$, not necessarily a solution of the equations of motion.
+That is, $d\vec{\gamma}$ will not be, in general, equal to $\vec{S}dt$.
 Consider the line integral of the vector potential $\vec{\theta}$ along $\gamma$.
@@
 	&= \int^{t_2}_{t_1} \left(p \frac{dq}{dt} + 0 \frac{dp}{dt} - H \frac{dt}{dt}\right) dt \\
-	&= \int^{t_2}_{t_1} \left(p \frac{dq}{dt} - H\right) dt
+	&= \int^{t_2}_{t_1} \left(p \frac{dq}{dt} - H\right) dt \\
+	&= \int^{t_2}_{t_1} L \, dt
 \end{aligned}
 \end{equation}
-In other words, the line integral of the vector potential over an evolution coincides with the action, and the Lagrangian is simply
-\begin{equation}
-	L = \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt}.
-\end{equation}
+Therefore the line integral of the vector potential over a path $\gamma$ coincides with the action $A[\gamma]$.
 
 \begin{figure}
-	\includegraphics[width = 0.45\textwidth]{ActionNonOptimized.png}
-	\includegraphics[width = 0.45\textwidth]{ActionOptimized.png}
+	a \includegraphics[width = 0.45\textwidth]{ActionNonOptimized.png} \\
+	b \includegraphics[width = 0.45\textwidth]{ActionOptimized.png}
 	\caption{\footnotesize{The variation of the action is the flow of the displacement field $\vec{S}$ through the surface $\Sigma$ that sits between the path $\gamma$ and its variation $\gamma'$.
@@
 
-Since the Lagrangian is the scalar product between the displacement and its vector potential, it suffers all the problems of the vector potential: it is not unique and its value is not directly physically meaningful.
+Since the Lagrangian is equal to the scalar product between the displacement along a path $\frac{d\vec{\gamma}}{dt}$ and the vector potential $\vec{\theta}$, it suffers all the problems of the vector potential: it is not unique and its value is not directly physically meaningful.
 The same is true for the action itself.
@@
 \begin{equation}
-	H = c |p|
+	H = c |p^i|
 \end{equation}
@@
 the trajectory of a photon is not enough to recover its full state).
-In these cases, the Lagrangian approach cannot be applied because $\vec{\theta} \cdot \vec{S}$ cannot be expressed as a function of position and velocity.
+In these cases, the Lagrangian approach cannot be applied because $\vec{\theta} \cdot \frac{d\vec{\gamma}}{dt}$ cannot be expressed as a function of position and velocity.
 Yet, if we express $L(q,p,t)$ in terms of conjugate momentum instead of velocity, the modified version of the variational principle still works.
@@
 
-\section{Multiple degree of freedom}
+\section{Multiple degrees of freedom}
 
 All our previous points will hold true in the generalization to multiple degrees of freedom.
-The only additional key insight is that determinism and reversibility will apply to each independent degree of freedom, meaning that we do not only preserve the total number of states of the system, but we preserve the fact that the total number of states is the product of the configurations given by each independent degree of freedom.
+The only additional key insight is that determinism and reversibility will apply to each independent degree of freedom, meaning that we do not only preserve the total number of states of the system, but also that the total number of states is the product of the configurations given by each independent degree of freedom.
 
 To express this requirement, the tools provided by differential calculus are not enough.
-We need elements of differential geometry, particularly symplectic geometry and contact manifolds, which is problematic for two reasons.
-First, these are not widely used (or understood) among physicists, and are not accessible to physics students.
-Second, the treatment of the subject is typically too abstract and uses notation that we find, in some cases, misleading to a physicist.
-This is unfortunate because the underlying concepts are quite accessible and provide a more general and robust foundations for many areas of physics, not just classical particle mechanics.
-
-Given that we want to make the generalization understandable to a wide array of people with physical background, we are going to use enough of the standard terminology from differential geometry that our methods are recognizable to someone proficient in it, but we are going to use a notation that parallel Einstein's notation, so that it feels more familiar to a physicist.\footnote{The main differences are as follows.
+We need elements of differential geometry, particularly symplectic geometry and contact manifolds.
+To make the treatment more appropriate for a physics audience, we are going to use a notation that parallels Einstein notation for general relativity.
+\footnote{The main differences are as follows.
 The vector basis will be noted as $e_a$ instead of $\frac{\partial}{\partial x^a}$.
@@
 Our system will be now composed of $N$ independent degrees of freedom.
-This means that we can chart all states at all times using $2N + $ variables $\xi^a = \{ q^i, p_i, t\}$.
+This means that we can chart all states at all times using $2N + 1$ variables $\xi^a = \{ q^i, p_i, t\}$.
 We will use $\xi^a$ when we want to span all variables; $q^i$ and $p_i$ when we want to span the position or the momentum of all degrees of freedom.
@@
 \end{equation}
-However, in this cases we will not use this object to fully characterize the geometry of the space.
+However, unlike in the single degree of freedom case, $\vec{S}$ is not the object used fully characterize the geometry of the space.
 
-The main idea is that we wont be able to quantify the number of states identified by each degree of freedom.
+The main idea is that we want to be able to quantify the number of states identified by each degree of freedom.
 As degrees of freedom are bi-dimensional, this means quantifying the areas of two-dimensional surfaces.
-Therefore we introduce the two-form
+Therefore we introduce the rank-2 tensor
 \begin{equation}
@@
 \end{equation}
-which we call the state counting form. The idea is that a pair of two vectors $\vec{v}$ and $\vec{w}$ will identify a parallelogram and 
+which we call the state counting tensor. The idea is that a pair of vectors $\vec{v}$ and $\vec{w}$ will identify a parallelogram in phase space and 
 \begin{equation}
@@
 \end{equation}
-quantifies the number of states over its surface. One can think of $\omega$ as a rank 2 tensor: a linear function of two vectors.
+quantifies the number of states over its surface.
 
-The form $\omega$ will need to satisfy additional condition to be physically meaningful.
+The tensor $\omega$ will need to satisfy additional conditions to be physically meaningful.
 First of all, it will need to be anti-symmetric.
@@
 \end{equation}
-This will mean the form will have $N(2N+1) = \sum_1^{2N}i$ independent components.
+Since the rank-2 tensor is anti-symmetric, it can be understood as a two-form in differential geometry.
+Only the components below the diagonal are independent, which means $\sum_1^{2N}i =N(2N+1)$ independent components.
 
@@
 over all closed surfaces.
-In fact, consider a parallelopiped.
+In fact, consider a parallelepiped.
 Given that the degrees of freedom are independent, and we have deterministic and reversible motion, two opposite sides will contain the same number of states.
 Given that the two opposite sides will have opposite orientation, their total contribution will be zero.
-This logic can be extended to all three-dimensional differential volumes, as they can be decomposed into infinitesimal parallelopipeds.
+This logic can be extended to all three-dimensional differential volumes, as they can be decomposed into infinitesimal parallelepiped.
 
@@
 \end{equation}
+The minus sign is the convention in symplectic geometry.
 
 In the generalization we need to distinguish between vectors, like $\vec{S}$, covectors, like $\theta$, and two-forms, like $\omega$.
-In three dimension, to each parallelogram there is only one perpendicular direction, which allows us to blur the line between the object in the following way.
-Each covector $\theta(\vec{S})$ can be expressed as a scalar product $\vec{\theta} \cdot \vec{S}$ with a suitable vector $\vec{\theta}$.
-Each two-form $\omega(\vec{v}, \vec{w})$ can be expressed as the triple product $\vec{\omega} \cdot \vec{v} \times \vec{w}$.
-The generalization of the curl is no longer an operation between vectors, but an operation that takes a one-form and returns a two-form.
+The single degree of freedom used only three variables, therefore to each parallelogram in $\{q, p, t\}$ there is only one perpendicular direction, which allows us to blur the line between the objects in the following way:
+\begin{itemize}
+	\item for each covector $\theta$ we can find a vector $\vec{\theta}$ such that $\theta(\vec{S})=\vec{\theta} \cdot \vec{S}$;
+	\item for each two-form $\omega$ we can find a vector $\vec{\omega}$ such that $\omega(\vec{v}, \vec{w}) = \vec{\omega} \cdot \vec{v} \times \vec{w}$.
+\end{itemize}
+This is what allows us to think only in terms of vectors.
+This cannot be done in the general case.
+In particular, the generalization of the curl is no longer an operation between vectors, but an operation that takes a one-form and returns a two-form.
 
-Now we need to further characterize the $\omega$.
-Variables $q^i$ and $p_i$ form independent degrees of freedom.
-This means that only surfaces spanned by matching position and momentum will define actual states, while all other combination will define no state.
+Now we need to further characterize $\omega$.
+Each pair of variables $q^i$, $p_i$ forms an independent degree of freedom.
+This means that only surfaces spanned by matching position and momentum will define actual states, while all other combinations will not define any state.
 Therefore:
@@
 \end{equation}
-The first condition sets $N^2$ independent components while the last two set $N(N-1)/2$ independent components each. This totals $N(2N-1) = \sum_1^{2N-1}i$ independent components.
+Of the $N(2N+1)$ independent components given the anti-symmetry requirement, the first condition sets $N^2$ of these while the last two set $N(N-1)/2$ independent components each.
+This totals $N(2N-1) = \sum_1^{2N-1}i$ independent components.
 
@@
 That is, no states appear or disappear from any degree of freedom.
-The displacement field doesn't define any state, no matter which other vector is paired with.
+The displacement field doesn't define any state, no matter which other vector it is paired with.
 Mathematically, we say that the displacement kills the form
@@
 
-Those familiar with symplectic and contact geometry, will recognize that conditions \ref{canonical_conditions} mean that $q^i$ and $p_i$ are canonical coordinates and, by Darboux's theorem, we will be able to write the covector potential as
+Those familiar with symplectic and contact geometry will recognize that conditions \ref{canonical_conditions} mean that $q^i$ and $p_i$ are canonical coordinates and, by Darboux's theorem, we will be able to write the covector potential as
 \begin{equation}\label{mdof_potential_expression}
@@
 	\omega(e^{q^i}, e^{p_j}) &= (-\partial\theta)_{q^i p_j} = -(\partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i}) = \delta^i_j \\
-	\omega_{q^i q^j} &= (-\partial\theta)_{q^i q^j} = -(\partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i}) = 0 \\
+	\omega(e^{q^i}, e^{q^j}) &= (-\partial\theta)_{q^i q^j} = -(\partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i}) = 0 \\
 	\omega(e^{p_i}, e^{p_j}) &= (-\partial\theta)_{p_i p_j} = -(\partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i}) = 0
@@
 
-We can then use, again, our gauge freedom with a function that does not depend on the first two variables.
-And we can use to set $\theta_{p_2} = 0$.
+We can then use, again, our gauge freedom with a function that does not depend on the first two variables to set $\theta_{p_2} = 0$.
 And, with the same reasoning, we will be able to set $\theta_{q^2} = p_2$.
@@
 \end{align*}
-Which recovers Hamilton's equation for multiple degrees of freedom.
+which recovers Hamilton's equations for multiple degrees of freedom.
 
-Once again, we have found that the whole physical content is included in conditions \ref{Detrev_assumptions}.
+Once again, we have found that the whole physical content is included in condition \ref{Detrev_assumptions}.
 The geometry in the extended phase space is fully specified by the state counting form $\omega$.
-This does not give define lengths and angles, like the metric tensor does for a more standard geometric setting.
+This does not define lengths and angles, like the metric tensor does for a more standard geometric setting.
 It defines areas of two-dimensional surfaces in terms of states, and angles between two-dimensional surfaces in terms of the degree of independence between degrees of freedom.
-What the equations say, is simply that states move in time at the same rate (i.e.
-$\frac{dt}{dt} = 1$) with an deterministic and reversible flow (i.e.
+What the equations say is simply that states move in time at the same rate (i.e.
+$\frac{dt}{dt} = 1$) with a deterministic and reversible flow (i.e.
 no states are created or destroyed on any independent degree of freedom).
@@
 
-The Lagrangian is the potential covector applied to the displacement field. That is
+The Lagrangian is equal to the potential covector applied to the displacement along a generic path. That is
 \begin{equation}
 \begin{aligned}
-L &= \theta(\vec{S}) = \theta_a S^a \\
-&= \theta_{q^i} S^{q^i} + \theta_{p_i} S^{p_i} + \theta_{t} S^{t} \\
+L &= \theta\left(\frac{d\vec{\gamma}}{dt}\right) = \theta_a \frac{d\xi^a}{dt} \\
 &= p_i \frac{dq^i}{dt} + 0 \frac{dp_i}{dt} - H \frac{dt}{dt} \\
@@
 
-As before, we will consider the variation of the line integral $\delta \int_{\gamma} \theta(d\gamma)$, which can be expressed as
+As before, we will consider the variation of the line integral $\delta \int_{\gamma} \theta(d\vec{\gamma})$, which can be expressed as
 \begin{equation}
 	\begin{aligned}
-		\int_A^B \theta(d\gamma) &= \int^{t_1}_{t_0} \theta(\vec{S}) dt \\
+		\int_A^B \theta(d\vec{\gamma}) &= \int^{t_1}_{t_0} \theta\left(\frac{d\vec{\gamma}}{dt}\right) dt \\
 		&= \int^{t_1}_{t_0} \left(p_i \frac{dq^i}{dt} - H \right) dt
@@
 \end{equation}
-The original path $\gamma$ together with its variation $\gamma'$ form a closed line $\lambda$.
-We can then use Stoke's theorem to transform the line integral of $\vec{\theta}$ to a surface integral of its exterior derivative.
+The original path $\gamma$ together with its variation $\gamma'$ form a contour $\partial \Sigma$.
+We can then use Stokes' theorem to transform the line integral of $\theta$ to a surface integral of its exterior derivative.
 That is:
@@
 	\delta \int_{\gamma} \theta(d\vec{\gamma}) = 
-	&= \int_{\gamma} \theta(d\vec{\gamma}) - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma'} \\
-	&= \oint_{\lambda} \theta(d\vec{\lambda}) \\
-	&= \iint_{\Sigma} \partial \theta (\vec{\Sigma}) \\
-	&= - \iint_{\Sigma} \omega(d\vec{\Sigma}).
+	&= \int_{\gamma} \theta(d\vec{\gamma}) - \int_{\gamma'} \theta(d\vec{\gamma}') \\
+	&= \oint_{\partial \Sigma} \theta(d\vec{\gamma}) \\
+	&= \iint_{\Sigma} \partial \theta (d\Sigma) \\
+	&= - \iint_{\Sigma} \omega(d\Sigma).
 \end{align*}
-One side of each $d\vec{\Sigma}$ will be $d\gamma$, therefore $\omega(d\vec{\Sigma}) = \omega(d\vec{\gamma}, \vec{v}) = \omega(\vec{S}dt, d\lambda)$ where $d\lambda$ is the direction of the variation.
-Since \ref{mdof_displacement_kills}, we have
+At each point, one side of $d\Sigma$ will be $d\vec{\gamma}$, therefore $\omega(d\Sigma) = \omega(d\vec{\gamma}, d\vec{\lambda})$ where $d\vec{\lambda}$ is the direction of the variation.
+If $\gamma$ is an actual evolution, $d\vec{\gamma} = \vec{S} dt$ and, by \ref{mdof_displacement_kills}, we have
 \begin{equation}
 	\begin{aligned}
-	\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} &= - \iint_{\Sigma} \omega(d\vec{\Sigma}) \\
-	&= - \iint_{\Sigma} \omega(\vec{S}, d\lambda) dt \\
-	&= - \iint_{\Sigma} 0 dt = 0
+	\delta \int_{\gamma} \theta (d\vec{\gamma}) &= - \iint_{\Sigma} \omega(d\Sigma) \\
+	&= - \iint_{\Sigma} \omega(d\vec{\gamma}, d\vec{\lambda}) \\
+	&= - \iint_{\Sigma} \omega(\vec{S}, d\vec{\lambda}) dt \\
+	&= - \iint_{\Sigma} 0 = 0
 	\end{aligned}
@@
 
-The action principle, then, is a direct consequence of the geometry set by \ref{mdof_closed_form} and \ref{mdof_displacement_kills}, which is the generalization of the divergence-free nature displacement field $\vec{S}$.
+The action principle, then, is a direct consequence of the geometry set by \ref{mdof_closed_form} and \ref{mdof_displacement_kills}, which is the generalization of the divergence-free nature of the displacement field $\vec{S}$.
 Every evolution $\gamma$ is a field line of $\vec{S}$.
-Any infinitesimal surface $\Sigma$ enclosed by an evolution and its variation, then, will also be tangent to the displacement field $\vec{S}$ and therefore will contribute no states, $\omega(d\Sigma)$ will be zero.
+Any infinitesimal surface $d\Sigma$ enclosed by an evolution and its variation, then, will also be tangent to the displacement field $\vec{S}$ and therefore will contribute no states, $\omega(d\Sigma)$ will be zero.
 Given that $\omega$ is closed, the integral over $\Sigma$ can be expressed as the line integral of the covector potential $\theta$ over the contour of $\Sigma$.
@@
 The additional key assumption is that the degrees of freedom must be independent.
-If the degrees of freedom were not independent, then the number of states within a four-dimensional parellopiped would not be given by the count of states on the sides and the angle.
+If the degrees of freedom were not independent, then the number of states within a four-dimensional parallelepiped would not be given by the product of the count of states on the sides.
 We would need additional structure.
-Conversely, if the degrees of freedom would not remain independent, the displacement field $\vec{S}$ would only preserve the count of states of volumes, not over each degree of freedom, and therefore $\omega(\vec{S}, \cdot) \neq 0$.
+If the degrees of freedom started independent but didn't remain so, the displacement field $\vec{S}$ would only preserve the count of states of volumes, not over each degree of freedom, and therefore $\omega(\vec{S}, \cdot) \neq 0$.
 
-The comments on the lack of strict physicality for the values of the Lagrangian and the action remains unchanged.
+The comments on the lack of strict physicality for the values of the Lagrangian and the action remain unchanged.
 They both depend on a choice of gauge.
-The comment on the centrality of the extended phase space to understand the action principle is only reinforce.
+The comment on the centrality of the extended phase space to understand the action principle is only reinforced.
 Additionally, note that by changing $t$ to $ct$ expression \ref{mdof_potential_expression} becomes
@@
 
-We have given a geometrical interpretation of the action principle that is also fully physically motivated by the assumption of deterministic and reversible evolution.
+We have given a geometrical interpretation of the principle of stationary action that is also fully physically motivated by the assumption of deterministic and reversible evolution.
 The key insight is that a path $\gamma$ and a variation form a closed curve which encloses a surface $\Sigma$.
 If the path is an actual evolution for the system, all states flow along the surface $\Sigma$, not through it.
-Using Stoke's theorem, the zero flow condition through the surface $\Sigma$ is transformed to a zero integral condition on its boundary, on the path $\gamma$ and its variation, leading to the principle of least action.
+Using Stokes' theorem, the zero flow condition through the surface $\Sigma$ is transformed to a zero integral condition on its boundary (i.e.
+on the path $\gamma$ and its variation) leading to the principle of least action.
 
@@
 This makes the numerical values of those objects not strictly physical.
-We have also seen that the interpretation given also works for those Hamiltonian systems that do not have a corresponding Lagrangian, meaning that the most general expression for the action principle lies on the extended phase space formulation of classical mechanics.
-We have also seen that elements from symplectic and contact geometry are needed for the generalization and that they can be formulated in a language that is less abstract and closer to the physics.
+We have seen that the interpretation given also works for those Hamiltonian systems that do not have a corresponding Lagrangian, meaning that the most general expression for the action principle lies on the extended-phase-space Hamiltonian formulation of classical mechanics.
+We have seen that elements from symplectic and contact geometry are needed for the generalization to multiple degrees of freedom and that they can be formulated in a language that is less abstract and closer to physics conventions.
 

### 2022-02-25 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QuantumLogic: first pass from reviewers

@@
 
+\newcommand\change[1]{{\color{red!75!black} #1}}
+
 \begin{document}
@@
 \begin{abstract}
-It is generally believed that quantum mechanics entails a rejection of classical logic.
+\change{At the onset of quantum mechanics, it was speculated that the new theory would entail a rejection of classical logic.
 The main arguments to support this claim come from the non-commutativity of quantum observables, which allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions.
-In this essay we show that such arguments contain serious flaws.
-In particular, it is possible to keep the classical distributivity law in the context of quantum theory once the actual meaning of quantum propositions is unambiguously defined and properly taken into account.
-We moreover argue that the existence of quantum superpositions does not lead to the definition of a new rule for logical disjunctions.
-Furthermore, contrary to the received view in quantum logic, we explain that quantum mechanics can generate a distributive lattice of propositions.
-Finally, the remarkable formal similarities between the logical structures of classical and quantum mechanics are highlighted and discussed.
+While the quantum logic program is not as popular as it once was, the core question remains unsettled: what is the relationship between the logical structures of the two theories?
+In this essay we show that the original arguments promoting quantum logic contain serious flaws, and that quantum theory does satisfy the classical distributivity law once the full meaning of quantum propositions is properly taken into account.
+Moreover, we show that quantum mechanics can generate a distributive lattice of proposition, which, unlike the one of quantum logic, includes statements about expectation values which are of undoubtable physical interest.
+Lastly, we show that the lattice of statistical propositions in classical mechanics follows the same structure, yielding an analogue non-commutative sublattice of classical propositions.
+This shows that the purported difference between classical and quantum logic stems from a misconstructed parallel between the two theories.}
 \vspace{4mm}
@@
 A few years after the publication of this paper, the research in quantum logic saw a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct logical structures to understand and describe the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{Pitowsky:2006}, \cite{Giuntini:2018}, \cite{Svozil:2020}, \cite{Holik:2021}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}).
-In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography.
-Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
+\change{While the approach is still being actively pursued by a sized community, and despite the many interesting applications in the fields of quantum information, computation and cryptography that are being investigated to the present day, QL is no longer viewed as a solution to the foundational problems affecting quantum theory.\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
 Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory].
-This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).}, it remains an important discipline of contemporary physics.
-In this paper we will neither take a stance in the dispute concerning the ontological status of logic and its laws, nor we will consider recent developments of QL given that they are not strictly relevant to the point we are going to make.
+This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).} Nevertheless, while it may be reasonable for a physicist to simply discard the approach over practical considerations, such as the lack of a desired result, on philosophical grounds it begs the question: what is the relationship between the logical structures of classical and quantum mechanics?
+How are they different and how are they similar?
+What should we think about the core insights that lead to the development of QL?
+Were they substantially correct, or should they be rejected?
+Given the fundamental role that logic plays in philosophy, we feel that these questions still must be answered regardless of how one may feel towards QL, or we are bound to make the same mistakes.} In this paper we will neither take a stance in the dispute concerning the ontological status of logic and its laws, nor we will consider recent developments of QL given that they are not strictly relevant to the point we are going to make.
 Indeed, the present essay is focused on the logical structures common to both quantum and classical mechanics, following the spirit of the former discussion.\footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM.
@@
 \end{description}
-\noindent Because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true.
-Therefore $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$.
-Consequently $q_i \vee r_i = q'_i \vee r'_i$.
+\noindent \change{If $\rho_i$ is the (potentially) mixed state before the measurement, the expectation value for $y$-spin is given by $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar \langle \psi_y^+ | \rho_i | \psi_y^+ \rangle - \frac{1}{2} \hbar \langle \psi_y^- | \rho_i | \psi_y^- \rangle$.
+Therefore $q'_i$ is true if and only if $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar$.
+Because $\rho_i$ is positive definite, this can only happen if $\langle \psi_y^+ | \rho_i | \psi_y^+ \rangle = 1$ and $\langle \psi_y^- | \rho_i | \psi_y^- \rangle = 0$, which means $\rho_i = | \psi_y^+ \rangle \langle \psi_y^+ |$.
+This means that the expectation value for $y$-spin is $\frac{1}{2} \hbar$ if and only if the electron was prepared in a pure state of $y$-spin up.
+In other words, because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true.
+We have that $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$.} Consequently $q_i \vee r_i = q'_i \vee r'_i$.
 Note that $q'_i \vee r'_i$  is the proposition ``before the measurement, the expectation value for $y$-spin is $\pm \frac{1}{2} \hbar$''.
@@
 Therefore it is not the case that the quantum disjunction can be true even if none of the elements is true.
+\change{The overall point is that the incompatibility of the observables does indeed introduce a problem because it is not clear how to evaluate propositions that mix different variables.
+The use of equivalent propositions over expectation values helps us solve the ambiguities.}
 
@@
 Therefore $V$ is not a closed subspace.
-Our proposition, despite being physically interesting, does not correspond to a closed subspace but does correspond to a Borel set.
+Our proposition, despite being physically interesting, does not correspond to a closed subspace but does correspond to a Borel set.\footnote{\change{Note that we are not claiming that all Borel sets can be constructed in this way or that all Borel sets correspond to physically interesting propositions.
+First of all, it is not true: $\{ \psi \}$ and $\{ e^{\imath \theta} \psi\}$ are different Borel sets, yet they are not physically distinguishable since a difference in absolute phase is not physically relevant.
+This can be fixed by considering the Borel sets of the projective space.
+Second, to really examine the physicality of all Borel sets we need a ``general theory of experimental logic'', of the type provided by [cite kevin kelly] or [cite carcassi aidala], which would go well beyond the present discussion.
+With appropriate caveats (e.g.
+the Hilbert space must be separable), one can say that the Borel sets corresponds to propositions that can be associated to an experimental test (what [carcassi aidala] call theoretical statements) which may or may not terminate in any or all cases.
+We leave this discussion for another work.}}
 

### 2022-02-25 Articles/2020-QuantumLogic/Source/response1.tex
QuantumLogic: first pass from reviewers

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amsmath, amsthm, amssymb}
\usepackage{amsfonts}
\usepackage{MnSymbol}
\usepackage{mathtools}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{enumitem}

\newcommand\change[1]{{\color{red!75!black} #1}}

\begin{document}

TO THE EDITOR

thank you very much for having provided us
good quality referee reports for our paper. We appreciate that this is
not an easy task.

\section*{Referee 1}

\begin{quote}
The topic of this paper is the standing of "quantum logic" vis-á-vis "classical logic", and in particular whether quantum mechanics demands either a rejection of or adjustment to classical logic. The answer, in a nutshell, is "no". The burden of the paper is to explain exactly why the standard arguments to the opposite conclusion are incorrect, and incidentally how both distributive Boolean lattices of propositions can be defined in quantum mechanics and non-distributive, non-Boolean lattices of propositions can be defined in the theory of probability distributions over classical phase space. These observations serve to show how the non-distributivity is not connected either to quantum mechanics or to the use of complex (rather than real) numbers.

The paper is clear and decisive. There are sufficiently many citations to justify publications, for even though the star of quantum logic has faded it has not entirely gone out. It is worthwhile to have all of this laid out in a single, compact presentation.

I have only a couple of comments/ suggestions.

p. 2: The quotation from Holik seems quite reasonable. All that Holik claims is that the connectives he is using in the context of quantum theory are "analogues" to the classical truth-functional connectives, and immediately mentions that they cannot be taken to be the classical connectives. Unless Holik adds that the quantum ones are meant to replace or correct classical logic, the statement seems fine. It is certainly much better than the Stefanovich talk of "replacement".

p. 9: "More precisely, we are going to argue that (i) it is not necessarily the case that a quantum
disjunction must be true when neither of its members is". But surely that is attacking a straw man. No one every claimed that a "quantum disjunction" (join) must be true when the propositions joined fail to be true! For example, when they are both false.

\end{quote}

TODO

\begin{quote}
p. 1 1: "The solution is simple: while
the value of an observable is not always defined, its expectation value always is. The idea is
then to translate qi and ri into propositions about expectation values, which we already know
how to handle, and apply the standard rules. Consider the following statements:

q'i: "before the measurement, the expectation value for y-spin is ½ h.

r'i: "before the measurement, the expectation value for y-spin is -½ h.

Because the spin of the electron along any axis is bounded by ±1/2 h, the only way that q'I is
true is if qi is true. Therefore q'I and r'i are respectively equivalent to qi and ri."

I can't follow this claim. It is true that the only way that the expectation value can be, say, ½ h is if the system is in an eigenstate of y-spin with the value up, but one could certainly imagine a situation (according to a theory) in which the particle does indeed have y-spin up but the expectation value associated with the system is not +1/2 h. This could be true on a statistical understanding of the expectation value: even though this system has y-spin up, the expectation value is calculated with respect to an ensemble of systems, not all of which have y-spin up. So the inference certainly goes through in one direction, but whether it does in the other depends on how the expectation values are understood. But "equivalent" suggests equivalence in both directions. So I think more would have to be said here.
\end{quote}

To remove ambiguity, we expanded the example to take the most general case, that of a density matrix, and show that fixing the expectation value to 1/2 h means the density matrix must correspond to the pure state of spin up. That way it is clear what we mean by expectation, and we cover all possible cases that are present in quantum mechanics (i.e. there is no mixed state that could give the same result). It reads:

\change{If $\rho_i$ is the (potentially) mixed state before the measurement, the expectation value for $y$-spin is given by $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar \langle \psi_y^+ | \rho_i | \psi_y^+ \rangle - \frac{1}{2} \hbar \langle \psi_y^- | \rho_i | \psi_y^- \rangle$. Therefore $q'_i$ is true if and only if $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar$. Because $\rho_i$ is positive definite, this can only happen if $\langle \psi_y^+ | \rho_i | \psi_y^+ \rangle = 1$ and $\langle \psi_y^- | \rho_i | \psi_y^- \rangle = 0$, which means $\rho_i = | \psi_y^+ \rangle \langle \psi_y^+ |$. This means that the expectation value for $y$-spin is $\frac{1}{2} \hbar$ if and only if the electron was prepared in a pure state of $y$-spin up. In other words, because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true. We have that $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$.}

Note that the argument holds regardless of how one interprets $\rho$ (e.g. state of knowledge, fraction of repeated experiments, ...).

\begin{quote}
I can see why there is a desire to introduce claims about expectation values: that make the transition to the discussion of classical physics smoother. But I'm not sure that this is the best way to do it. The main point about being careful with the time indices is the key to resolving the non-distributivity issue, and that can be does without bringing in expectation values. So this passage might be made clearer.
\end{quote}
Actually, the smoother transition is a bonus. The main reason to introduce expectation values here is to settle the question of how to evaluate statements of over one observable when an incompatible observable is prepared. As the text reads ``The difficulty here is that before the measurement of $S_y$ only $x$-spin is defined, so how are we going to compute statements about the value of $y$-spin?'' We suspect that, given the length of the paragraph, the point is lost by the end, so we close the paragraph with:

\change{The overall point is that the incompatibility of the observables does indeed introduce a problem because it is not clear how to evaluate propositions that mix different variables. The use of equivalent propositions over expectation values helps us solve the ambiguities.}

So the idea is that it is useful to add these propositions anyway, so it already gives us a hint to compare to classical statistical propositions.

\section*{Referee 2}

\begin{quote}
This is a very interesting article, and it puts back on the table an older issue that had been dormant for a while. As such, it is a helpful contribution. However, it does have several shortcomings that prevent me from endorsing for publication in its present form. 
\end{quote}

\begin{quote}
1. The primary defect of the paper is that it greatly overstates the degree to which people are convinced that QM demands a revision of our logic. I would say that 90\% of philosophers of physics believe that it does *not* demand a revision of our logic, and that 90\% of physicists would say that they do not know what it means to say that QM demands a revision of logic. The citations that the authors use represent only a small minority of philosophers and physicists who were trying to sell this idea.

The main problem here is not that the authors misdescribe the situation. The main problem is that their arguments are largely aimed at a straw person -- and hence it is not so likely that the paper would receive a broad readership. 
\end{quote}

Looking again at the written prose, the stated motivation of the paper (particularly the abstract) is indeed just an argument against Quantum Logic. In that sense, the referee is correct. On one hand, QL still attracts funding and interest and, as referee 1 puts it:
\begin{quote}
	... for even though the star of quantum logic has faded it has not entirely gone out. It is worthwhile to have all of this laid out in a single, compact presentation.
\end{quote}
So we still think talking about QL is relevant, However, as the referee 2 notes later, the article goes past QL and compares at a deeper level the logical structures of classical and quantum mechanics. This, we believe, is of general interest regardless of how one may feel about QL. Given  comments from both referees, we propose to keep the arguments against QL, but to make it clear that our discussion goes deeper.

We revised the abstract as follows:

\change{At the onset of quantum mechanics, it was speculated that the new theory would entail a rejection of classical logic. The main arguments to support this claim come from the non-commutativity of quantum observables, which allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions. While the quantum logic program is not as popular as it once was, the core question remains unsettled: what is the relationship between the logical structures of the two theories? In this essay we show that the original arguments promoting quantum logic contain serious flaws, and that quantum theory does satisfy the classical distributivity law once the full meaning of quantum propositions is properly taken into account. Moreover, we show that quantum mechanics can generate a distributive lattice of proposition, which, unlike the one of quantum logic, includes statements about expectation values which are of undoubtable physical interest. Lastly, we show that the lattice of statistical propositions in classical mechanics follows the same structure, yielding an analogue non-commutative sublattice of classical propositions. This shows that the purported difference between classical and quantum logic stems from a misconstructed parallel between the two theories.}

We revised part of the introduction as follows:

\change{While the approach is still being actively pursued by a sized community, and despite the many interesting applications in the fields of quantum information, computation and cryptography that are being investigated to the present day, QL is no longer viewed as a solution to the foundational problems affecting quantum theory.\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}. Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory]. This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).} Nevertheless, while it may be reasonable for a physicist to simply discard the approach over practical considerations, such as the lack of a desired result, on philosophical grounds it begs the question: what is the relationship between the logical structures of classical and quantum mechanics? How are they different and how are they similar? What should we think about the core insights that lead to the development of QL? Were they substantially correct, or should they be rejected? Given the fundamental role that logic plays in philosophy, we feel that these questions still must be answered regardless of how one may feel towards QL, or we are bound to make the same mistakes.}

TODO: Maybe review the conclusions as well?

\begin{quote}
2. In Section 3, the authors argue the logic of quantum propositions is distributive ... if one indexes the propositions so that they refer to times (either before or after measurement). This point is not novel. It was already pointed out (although vaguely) by Niels Bohr in reply to John von Neumann. It was again pointed out by Patrick Heelan in the 1970s, and by Stanley Gudder shortly thereafter. The point of all these suggestions is roughly that if one indexes propositions by their context, then a non-distributive logic can be represented as a contextual distributive logic. 
\end{quote}

\begin{quote}
3. Section 4 is actually innovative, and could well be developed into a novel point. Especially interesting is the argument (pages 16-18) that all Borel subsets of Hilbert space represent physically interesting propositions about the system. But let me note a few qualifications. (a) The authors do not show that *every* Borel subset of H is of the form described in Proposition 2. That would be a very interesting result. (b) The authors do not cite others who have argued that all Borel subsets of H are physically relevant. See especially David Wallace's book on the quantum multiverse. (c) The authors do not consider any of the many objections that could be raised against their proposal. For example, one would not normally say that "quantity A having expectation value r" is a property of a system, but that it is a meta-property representing our ignorance of the actual value of A.
\end{quote}

For (a), the situation is a bit more complicated. Let us mention a few problems.

Not all Borel subsets of H are of the form described in Proposition 2. In fact, they are not all physically relevant. The issue is that a state is not a vector in the Hilbert space, but a ray (i.e. a subspace of dimension 1). The absolute phase and absolute norm are not physically observable per se, though these can be differentiated with Borel sets. The proper physically meaningful state space would be the projective space, whose Borel sets are a subset of the Borel sets of the Hilbert space.

Additionally: not all rays can realistically be considered physical. Some states, for example, correspond to infinite expectation for position, momentum, energy. One can, however, show that the set of states with finite expectation values for all polynomial of position and momentum corresponds to the Schwartz space, which is dense in the Hilbert space. One can then show that a basis for the topology of the projective space of the Schwartz space can be constructed with sets of the form described in Proposition 2.

More in general, to make a solid map between statements and experimental verification, the best way would be to use the results of Kelly or Carcassi/Aidala that link the open set of a topology with experimentally verifiable. This would also allow to go in further details. For example, for the space to be (at least in the limit) explorable experimentally the topology must be second countable and $T_0$. Outside of those restriction, one can show that the points of the space do not correspond to cases that can be experimentally distinguished.

Note that all these issues apply in an analogous way for the lattice of classical statistical propositions we consider.

The point here is that we would need the appropriate space to go through all of these details, which would probably double the size of the paper as we would have to introduce a lot of previous work that we cannot expect the reader to be familiar with. Our purpose therefore is limited to only show that the Borel sets contain a big class of experimentally interesting propositions that are not contained in the QL logic lattice.

We added the following footnote to both clarify and give some more information about the issue:

\change{Note that we are not claiming that all Borel sets can be constructed in this way or that all Borel sets correspond to physically interesting propositions. First of all, it is not true: $\{ \psi \}$ and $\{ e^{\imath \theta} \psi\}$ are different Borel sets, yet they are not physically distinguishable since a difference in absolute phase is not physically relevant. This can be fixed by considering the Borel sets of the projective space. Second, to really examine the physicality of all Borel sets we need a ``general theory of experimental logic'', of the type provided by [cite kevin kelly] or [cite carcassi aidala], which would go well beyond the present discussion. With appropriate caveats (e.g. the Hilbert space must be separable), one can say that the Borel sets corresponds to propositions that can be associated to an experimental test (what [carcassi aidala] call theoretical statements) which may or may not terminate in any or all cases. We leave this discussion for another work.}

For (b), TODO

For (c), each interpretation of quantum mechanics will have a competing view on the meaning of the expectation values and on the meaning of the propositions. For example, for some (e.g. Qubism, TODO) all expectations will indeed provide only information about the state of knowledge of the agent because, in their view, that is all that quantum states provide. For other (e.g. Bohmian, TODO) they will represent expectation over hidden variables.

Regardless of the interpretation, the framework of quantum mechanics provides a well-defined expectation value operation for either a  pure state $\psi$ and a density matrix $\rho$. Therefore the statement ``quantity A has expectation value r'' will correspond to the mathematical expression $\langle \psi | A | \psi \rangle = r$ (or $\text{Tr}(\rho) = r$). It is up to the particular interpretation to make it clear what those expectation mean. The logical structure (i.e. the Borel sets) will be the same regardless.

TODO: add a footnote somewhere?

--------------------

In closing, we were very pleasantly surprised by the quality of both
reports and we thank both Referees for the time they spent in
evaluating our paper.

Sincerely,

The Authors

\end{document}

### 2022-03-05 Articles/2021-ReversePhysics/Submissions/FoP/ReversePhysicsSupplementary.tex
Materials corresponding to FoP submission

--- 2021-11-08 Articles/2021-ReversePhysics/ReversePhysics.tex
@@
 
-\title{Reverse Physics}
+\title{Reverse Physics: Supplementary Information}
 \author{Gabriele Carcassi, Christine A. Aidala}
@@
 
-\begin{abstract}
-We introduce an approach to the foundations of physics that is more in line with the foundations of mathematics.
-The idea is to examine current theories and find a set of starting physical assumptions that are sufficient to rederive them, which is the physical equivalent of the reverse mathematics program.
-This allows us to identify physical starting points (as opposed to mathematical ones), compare their deductive power, and find those that can better serve as a foundation for our theories.
-\end{abstract}
-
 \maketitle
 
-\section{Introduction}
-
-%\begin{itemize}
-%	\item Brief descriptions of the aims of foundations of mathematics
-%	\item We want to follow a similar approach (not new interpretations of QM or new models for quantum gravity)
-%	\item Reverse mathematics start from theorems and finds the smallest set of axioms needed to rederive them.
-We want to take physical theories and find the smalles set of physical assumptions that are equivalent to the theories themselves.
-%	\item This paper is meant to showcase the approach and how it can lead to renewed physical insights into theories that may be thought as already fully understood and those that are not.
-The main point of the paper is not the results per se, but the novel of the approach proposed.
-We hope that other people will see the value and pursue similar lines of investigations.
-%	
-%\end{itemize}
-
-Within the foundations of mathematics, the reverse mathematics program\cite{friedman1976systems,simpson2017reverse,stillwellreverse} aims to start from known theorems and determine what axioms are required to prove them.
-This helps understand the relative logical strength of the theorems and get a better sense of their relationships.
-We propose a similar program for physics.
-We start from known physical theories or elements of physical theories, and we try to find premises that enable us to rederive them.
-The goal is to find premises that are physically more intuitive, so that ideally the arguments can be carried out with or without the math, and develop concepts that are of physical significance across theories.
-
-While one does find efforts that analyze the premises of physical theories (see e.g.
-Refs.~\cite{chiribella2011informational, selby2021reconstructing, giles2016mathematical, boyling1972axiomatic, PhysRev.101.860, Haag1964848, axiomaticQFT1975, masanes2019measurement, carcassi2021four}), for example in the context of foundations of quantum mechanics, statistical mechanics or the search for quantum gravity, their aim is typically different.
-Some are focused on a particular field (e.g.
-quantum theory) or try to demonstrate the centrality of one aspect (e.g.
-information) with regard to others.
-Some concentrate more on mathematical premises or structure, making the physical significance less than clear.
-Many aim to develop new theories, instead of first gaining further insight within current ones.
-We propose a more generalist approach that examines the relationships of physical ideas across subfields, which is best explained through examples.
-
-To give a sense of the breadth of the reverse physics approach, we show how it can be used to pursue three different objectives.
-First, we illustrate multiple instances of \emph{reformulation}.
-We study Hamilton's equations for a single degree of freedom to find six alternative characterizations that will connect different ideas from different fields under the same theme of determinism and reversibility.
-Next, we illustrate \emph{dependence analysis}.
-We show that the uncertainty principle is not a consequence of the full quantum theory, but simply of its zero entropy bound on pure states.
-If a similar bound is imposed in classical mechanics, similar uncertainty relationships are recovered.
-Lastly, we illustrate \emph{reconceptualization}.
-We take the third law of thermodynamics and rework it so that the role of crystalline substances to define zero entropy is instead played by the ``empty system''.
-
-As we want the discussion to be focused on the physics, we will use the most widely known math among physicists and relegate mathematical derivations to the appendix.
-We also want to stress that this paper is not about the originality of each single result, but about how these are used to address a more general and higher goal: to uncover the core physical concepts and assumptions in the known physical theories and understand their generative power and their interrelationships.
-
-%Use "investigation trumps rigor"
-%Somewhere, note that "physics" entropy does not coincide with "information" entropy. Physics entropy is information about states as quantified by invariants.
-
-\section{Determinism, reversibility and Hamiltonian mechanics}
-
-%\begin{itemize}
-%	\item Start with the equations of HM
-%	\item Show Liouville's theorem and wonder what it means. Jacobian is 1. Show how Jacobian for 1 d.o.f. is really a Poisson bracket.
-%	\item Connect to statistical mechanics: volume is number of states. HM conserves the number of states. States are mapped one to one on a continuum. (discuss problems of counting on a continuum)
-%	\item Show that we can go backwards on a single degree of freedom
-%	\item Wonder what to do on multiple d.o.f. 
-%	\item Show symplectic form as the geometrical description. Generalize to d.o.f. by taking outer product. Show that areas remain orthogonal
-%	\item What does orthogonality mean? Independence of d.o.f. (volume product of areas)
-%	\item Show that preservation of the geometry (areas and perpendicularity) under time evolution gives us HM
-%	\item HM is det/rev motion of indepent d.o.f.
-%	\item We used count of states, log count of states is also conserved. What about information entropy? It is also conserved. 
-%	\item Show we can proceed the other way. Det/rev means information before and after is the same. Information needs to be conserved -> Jacobian = 1 -> HM.
-%	\item Note how bijection is not enough. There is an additional constraint. Why?
-%	\item Measurements on a continuum have uncertainty.
-Connect to statistics and use the covariance matrix to describe a gaussian peak.
-The determinant of the matrix with the Jacobian.
-If jacobian is 1, determinant is the same, same precision of the description.
-HM <-> to requiring a bijection at the same level of precision.
-%	\item Wonder why we are getting conservation of energy from det/rev, while we typically associate conservation of energy with system isolation (in the thermodynamic sense).
-Argue that if a system is deterministic and reversible, it means that the future/past state depends only on the system itself.
-The system does not depend on anything else -> it is isolated.
-There is a deep connection between isolation and det/rev
-%	
-%\end{itemize}
-
-Let us start our discussion with the equations of motion given by classical Hamiltonian mechanics:
-\begin{align}\label{HamiltonEquations1}
-	\frac{dq}{dt} = \frac{\partial H}{\partial p} \; \; \; \; \; \; \frac{dp}{dt} = - \frac{\partial H}{\partial q}
-\end{align}
-Our first task is to find a set of mathematical conditions that are equivalent to these equations.
-
-Let us group the state variables $\xi^a = \{q, p\}$ and consider the displacement vector field $S = \frac{d\xi^a}{dt} = \{ \frac{dq}{dt}, \frac{dp}{dt} \} $ which tells the direction in phase space in which each state moves over time.
-We find that $S$ is divergenceless:
-\begin{align}\label{divergenceless}
-\frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} =0
-\end{align}
-meaning that the flow through any closed region in phase space is zero. We also find the converse is true: a two dimensional divergenceless field $S$ always allows a stream function $H$ such that
-\begin{align}
-	S^q = \frac{\partial H}{\partial p} \; \; \; \; \; \; S^p = - \frac{\partial H}{\partial q}
-\end{align}
-In other words, Hamiltonian evolution for one degree of freedom is exactly the evolution for which the flow over a closed region in phase space is zero.
-
-Let us now look at how an infinitesimal region of phase space evolves.
-Suppose $Q = q + \frac{dq}{dt} dt = q + S^q dt$ and $P = p + \frac{dp}{dt} dt = p + S^p dt$.
-The infinitesimal region of area $dq dp$ will become $dQ dP = |J| dq dp$ where:
-\begin{align}
-	|J| = \begin{vmatrix}
-		\frac{\partial Q}{\partial q} & \frac{\partial Q}{\partial p} \\
-		\frac{\partial P}{\partial q} & \frac{\partial P}{\partial  p} 
-	\end{vmatrix} = \frac{\partial Q}{\partial q} \frac{\partial P}{\partial  p} - \frac{\partial P}{\partial q} \frac{\partial Q}{\partial p}
-\end{align}
-Note how the Jacobian determinant, in this simple case, coincides with the Poisson bracket. With a simple substitution we find:
-\begin{align}
-	|J|\approx 1 + \left( \frac{\partial S^q}{\partial q} + \frac{\partial S^p}{\partial p} \right)dt.
-\end{align}
-Note that the first order term is exactly the divergence, therefore condition \eqref{divergenceless} is equivalent to
-\begin{align}\label{UnitaryJacobian}
-	|J| = 1
-\end{align}
-which means the initial area $dqdp$ is equal to the final area $dQdP$.
-
-So we have found that, for a two dimensional manifold, Hamiltonian evolution \eqref{HamiltonEquations1}, zero flow over closed regions \eqref{divergenceless} and preservation of area \eqref{UnitaryJacobian} are mathematically the same condition.\footnote{This is essentially a short proof for Liouville's theorem that can work in both directions.} However, we still have to answer the main question: what does this mean physically?
-
-To do that, we look at statistical mechanics.
-There we use the area of a phase space region to count the number of states.
-Area preservation, then, means that the evolution preserves the state count: we start and we end with the same number of states.
-We have a one-to-one map between past and future states.
-Given an initial state we have one and only one final state.
-Physically, this means
-\begin{align}\label{DeterminismReversibility}
-	\text{Deterministic and reversible evolution}
-\end{align}
-is equivalent to conditions \eqref{HamiltonEquations1}, \eqref{divergenceless} and \eqref{UnitaryJacobian}.
-But determinism and reversibility is a physical condition, a physical property of the evolution: we have a physical characterization of Hamiltonian evolution.
-
-This seems to tell us that the physical requirement of deterministic and reversible evolution over the continuum mathematically is not simply a bijection.
-The reason is that, over a continuum, counting points is not enough.
-A segment a meter long has as many points as a segment a kilometer long, but they are not the same length.
-In the same way, all finite regions of phase space have infinitely many points, but physically they do not contain the same number of states.
-Mathematically, we need a measure to give a size to each region, and therefore deterministic and reversible evolution is a bijection that preserves the measure, that preserves the state count: Hamiltonian evolution.
-
-Note that this makes sense from a statistical mechanics/thermodynamics perspective as well.
-In statistical mechanics, the entropy is the logarithm of the count of states, therefore conservation of number of states means conservation of entropy.
-This means a deterministic evolution that is thermodynamically reversible (conservation of entropy) is also reversible in the dynamical sense (conservation of state count).
-Therefore
-\begin{align}\label{DeterminismThermodynamicReversibility}
-	\parbox{2.8in}{Deterministic and thermodynamically reversible evolution}
-\end{align}
-is yet another equivalent condition.
-
-What about entropy in the information sense?
-If the evolution is deterministic and reversible, the amount of information needed to specify the initial state should be exactly the same as the amount of information needed to specify the final state.
-That is, if $I[\rho] = - \int \rho \log \rho dqdp$ is the Gibbs/Shannon entropy, we would expect the following condition:
-\begin{align}\label{EntropyConservation}
-	I[\rho(t_1)] = I[\rho(t_2)]
-\end{align}
-to be equivalent to the others. We find:
-\begin{align}
-	I[\rho(t + dt)] = I[\rho(t)] - \int \rho \log |J| dqdp.
-\end{align}
-Since the Jacobian determinant of a continuous transformation cannot be negative, condition \eqref{EntropyConservation} is indeed equivalent to \eqref{UnitaryJacobian} and therefore to all the others.
-
-It may seem odd that conservation of entropy for a deterministic system gives us energy conservation: we typically associate energy conservation with system isolation in thermodynamics.
-Where is the connection?
-For a deterministic and reversible system, future and past states depend only on the state of the present system.
-Therefore they do not depend on the state of other systems or of the environment.
-This means that deterministic and reversible systems are necessarily isolated.
-This time, running the argument in the opposite direction presents a problem.
-One needs to first decide whether a non-deterministic isolated system makes sense.
-If one assumes that the source of non-determinism, of stochastic uncertainty, is always interaction with other systems, other degrees of freedom, then every isolated system is deterministic.
-In this case, we can run the argument in reverse: every isolated system is deterministic and reversible.
-Without that extra assumption, this reverse direction is not guaranteed.
-
-Finally, let us go back to the idea that a bijection is not enough to characterize deterministic and reversible evolution.
-Preservation of a measure is a mathematical idea.
-Is there a more physical way to look at it?
-The issue is that scientific measurements over a continuum can only carry finite precision.
-When considering how the information goes back and forth in time, then, we also need to take into account how the finite precision is mapped.
-While it may be true that the evolution of a damped harmonic oscillator is a bijection, the points get denser and denser around the equilibrium.
-Once we are close to equilibrium it becomes impossible to tell when the oscillator was started: the damped harmonic oscillator is not reversible in any practical sense if precision is taken into consideration.
-Let us see if this argument can be run formally.
-We want a coordinate invariant quantity that characterizes uncertainty.
-The obvious choice is the determinant of the covariance matrix:
-\begin{align}\label{CovarianceMatrix}
-	|\Sigma| = \begin{vmatrix}
-		\sigma_q^2 & cov_{q,p} \\
-		cov_{p,q} & \sigma_p^2
-	\end{vmatrix} = \sigma_q^2 \sigma_p^2 - cov_{q,p}^2
-\end{align}
-We can therefore imagine the following condition
-\begin{align}\label{BijectionUncertainty}
-	|\Sigma(t_1)| = |\Sigma(t_2)| 
-\end{align}
-for which the uncertainty remains constant in time. Is this yet another equivalent condition? If we assume the spread of the distribution is small enough, we find
-\begin{align}
-	|\Sigma(t + dt)| = |J| |\Sigma(t)| |J|.
-\end{align}
-therefore \eqref{BijectionUncertainty} is indeed equivalent to \eqref{UnitaryJacobian} and to all others.
-
-We could go on to find other relationships and extend the ones we found to multiple degrees of freedom, but we believe this should be enough to highlight the power of the reverse physics approach.
-The first thing to note is how in a couple of pages we have found fundamental connections between classical Hamiltonian mechanics \eqref{HamiltonEquations1}, vector calculus \eqref{divergenceless}, differential geometry \eqref{UnitaryJacobian}, statistical mechanics \eqref{DeterminismReversibility}, thermodynamics \eqref{DeterminismThermodynamicReversibility}, information theory \eqref{EntropyConservation} and plain statistics \eqref{BijectionUncertainty}.
-This helps foster a sense of unity of these disparate disciplines and their perspectives, a sense that is sorely lacking both in research and education: nature is one, and does not care about how we have divided academic knowledge.
-We believe that a single unified view of this kind will bring more coherence and clarity to physics than, for example, a unified theory for the fundamental forces.
-
-Another advantage is that we were able to carry out many of the arguments conceptually.
-Mathematics is then used to better articulate the physical arguments.
-This should please those who believe that physics, not mathematics, should be more at the center of the discussion.\cite{hossenfelder2018lost, woit2006not}
-
-The other interesting aspect is how much we were able to find in theories that are generally considered well understood.
-While it is common for some people to be aware of some of these results, we have found that most results are unknown to most people.
-
-\section{Uncertainty principle revisited}
-
-%\begin{itemize}
-%	\item We saw that classical mechanics is more than just points: it says something about uncertainty as well. Quantum mechanics has an uncertainty principle. Is there any connections?
-%	\item Since there was also a connection to entropy, turn to entropy in quantum mechanics.
-Entropy is bounded from below: all pure states have zero entropy.
-In classical mechanics, however, we can prepare distributions with arbitrarily small entropy.
-Can the fact that the entropy is bound from below in QM be the source of the uncertainty?
-%	\item Assume entropy is bounded from below in CM. Show you get uncertainty relationship in a single degree of freedom.
-%	\item Uncertainty relationship <-> bounded entropy = bounded information
-%\end{itemize}
-
-Let us now turn our attention to quantum mechanics, and see if our approach can shed new light on a theory that is not generally considered to be well understood.
-Since we talked about the role of uncertainty in classical mechanics, let us concentrate on the uncertainty principle, which states that every state has to satisfy the relationship
-\begin{align}\label{UncertaintyPrinciple}
-	\sigma_q \sigma_p \geq \frac{\hbar}{2}.
-\end{align}
-
-We can start with an interesting observation. If we look back at \eqref{CovarianceMatrix}, we have:
-\begin{align}
-	\sigma_q^2 \sigma_p^2 = |\Sigma| + cov_{q,p}^2 \geq |\Sigma|.
-\end{align}
-Since by \eqref{BijectionUncertainty} $|\Sigma|$ is a constant of motion, we find that during Hamiltonian evolution the uncertainty is bounded.
-The lowest uncertainty is reached when there are no correlations and the covariance is zero.
-Though it is a conceptually different relation, since $|\Sigma|$ is just a constant of motion, it gives us the following intuition: a coordinate invariant cap to the uncertainty will produce an inequality on the product of variances.
-Where can we find such a cap in quantum mechanics?
-
-Let us turn our attention to entropy, represented by the von Neumann entropy $I[\rho] = - tr (\rho \log \rho)$. For every pure state $|\psi \rangle$ we have the following property:
-\begin{align}\label{ZeroQuantumEntropy}
-	\begin{split}
-		I[|\psi\rangle\langle\psi|] = 0.
-	\end{split}
-\end{align}
-In other words, all pure states have zero entropy, the entropy is capped. Since entropy and uncertainty are related, is fixing the amount of entropy enough to recover the uncertainty principle?
-
-We can test that hypothesis by studying the space of classical distributions with a fixed value of entropy $I_0$.
-We find that they have to satisfy the relationship\footnote{The careful reader will note the units do not quite work.
-The issue here is the $\log \rho$ in the Shannon/Gibbs entropy, as $\rho$ is not a pure number.
-Introducing a dimensionful constant $\log \rho \hat{h}$ would fix the expression, which would then fix the uncertainty relationship as well.
-}
-\begin{align}\label{ClassicalUncertaintyPrinciple}
-	\sigma_q \sigma_p \geq \frac{e^{I_0}}{2 \pi e}.
-\end{align}
-The equality holds if the distribution is the product of two independent Gaussians.
-
-The following condition:
-\begin{align}\label{BoundedEntropy}
-	\text{The entropy of the system is fixed}
-\end{align}
-is therefore enough to recover an uncertainty relationship independently of the theory.
-This can be understood as fixing the amount of information carried by the state.
-The reverse argument does not work, so condition \eqref{BoundedEntropy} is a stronger condition than the existence of an uncertainty relationship.
-
-This example shows another advantage of reverse physics: it allows us to pinpoint what part of a theory is responsible for which effect.
-In this case, we have found that it is not quantum mechanics per se that leads to uncertainty relationships, just the entropy cap on pure states.
-If we implemented a similar cap in classical mechanics, we would obtain the same result.
-Therefore condition \eqref{ZeroQuantumEntropy} is a more fundamental and clear way to characterize the uncertainty in quantum mechanics, making it evident that all pure states carry the same information, they describe the system at the same level of precision.
-
-\section{A new understanding of the third law}
-
-%\begin{itemize}
-%	\item We saw that in quantum mechanics entropy is bounded from below. Why?
-%	\item Turn to thermodynamics.
-The third also says that thermodynamic entropy (not information entropy) is bounded from below.
-The zero thermodynamic entropy, however, is formulated in terms of a pure crystal at zero temperature.
-This does not seem to be a fundamental characterization.
-Can we do better?
-%	\item Note how entropy is an extensive quantity.
-Now consider the empty system.
-Since the composition of any system with the empty system must yield the same system, it also must yield the same entropy.
--> Entropy of empty system must be zero.
-Better more fundamental characterization of zero entropy.
-%	\item What would a system with negative entropy correspond to?
-To a description of a system that is more precise than saying the system is not there.
-But that is absurd!
-Therefore, it is a logical necessity that the entropy of a system is bound from below.
-%\end{itemize}
-
-In the previous section we saw the lower entropy bound of zero entropy is built into quantum mechanics and not in classical mechanics.
-However, the bound is also built into thermodynamics through the third law\footnote{This formulation is given by Ref.
-\cite{lewis1923thermodynamics}}
-\begin{align}\label{ThirdLaw}
-	\parbox{2.8in}{Every substance has a finite positive entropy, but
-		at the absolute zero of temperature the entropy may become zero, and does so
-		become in the case of perfect crystalline substances.}
-\end{align}
-Therefore we have two fundamental physical theories that share this trait. Can we find a more general characterization for this?
-
-The first step is to look for a system that can conceptually function as a zero for entropy that feels a bit less arbitrary than a perfect crystalline substance at zero temperature.
-Let us recall that a fundamental property for thermodynamic entropy is that it is an extensive quantity, it is additive for independent systems:
-\begin{align}
-	I_{AB} = I_{A} + I_{B}.
-\end{align}
-Is there a system that acts as a zero for system composition?
-This would be the empty system $\emptyset$: any system combined with the empty system $\emptyset$ will remain unchanged.\footnote{If we consider systems as a monoid under composition, the empty system is the identity element, much like the number zero, the empty set or the identity map in their respective structures.} In particular, we must have:
-\begin{align}\label{EntropyEmptySystem}
-	I_{\emptyset} = I_{\emptyset\emptyset} = I_{\emptyset} + I_{\emptyset} = 0.
-\end{align}
-We can therefore use the empty system as our zero reference for entropy.
-
-This move is a conceptual shift.
-In thermodynamics, the idea that the entropy for crystalline substances is zero is a phenomenological assumption: it is so because it justifies the behavior of substances as they approach absolute zero.
-We need statistical mechanics to explain it.
-Condition $\eqref{EntropyEmptySystem}$, however, is not a phenomenological consideration but a conceptual necessity: it must be so.
-In this sense, the new zero entropy system is more fundamental.
-But does it relate to the old ideas?
-In terms of statistical mechanics, the crystalline substance has zero entropy because it has one possible way to be.
-The empty system also has one possible configuration.
-So the old ideas indeed carry over: a crystalline substance will have to have the same entropy as the empty system.
-
-Now that we have a better concept for zero entropy, we need to understand why we can't have states with lower entropy.
-Let us consider it from an information theoretical perspective.
-Suppose we had a state with negative entropy.
-This would mean that it is better specified than a state with zero entropy.
-The state is better specified than saying that the system is not there.
-But stating that the system is not there already completely specifies the system.
-Therefore the condition that imposes a lower bound of zero for entropy is the following:
-\begin{align}\label{NewThirdLaw}
-	\parbox{2.8in}{No state can describe a system more accurately than stating the system is not there in the first place.}
-\end{align}
-Again, note that this condition is not phenomenological: it is a logical necessity.
-This condition is realized in quantum mechanics: pure states represent the most precise descriptions of the system and they have zero entropy much like the vacuum, another pure state.
-This condition is realized in thermodynamics through the third law.
-In fact, the third law could be rederived as follows.
-Assuming \eqref{NewThirdLaw} tells us that entropy is bounded at zero.
-Since the entropy in thermodynamics is a convex function of energy, entropy will reach its minimum at zero temperature, which is the only temperature that can reach zero entropy.
-We then use the previous result that the entropy of a crystalline substance at zero temperature is the same as for an empty system.
-
-This showcases the final advantage of reverse physics.
-We want to differentiate between ``assumptions'' (conditions that are not necessarily true) and ``principles'' (those that have to be taken to be true); we want to differentiate between conditions that are ``phenomenological'' (empirically derived or justified) from those that are ``conceptual'' (logical necessities or consequences of definitions).\footnote{Philosophers may call these ``constitutive'' conditions.} If we are able to elevate assumptions to principles, or replace phenomenological assumptions with conceptual ones, then the premises for our theories are sturdier and we have improved their foundations.
-
-%\section{A more unified physics}
-%
-%\begin{itemize}
-%	\item Put physics back at the center of the discussion (we are exploring physical concepts and their relationships, the math is there to carry out the arguments)
-%	\item Appeal to tradition (thinking about small set of physical assumptions was what foundations of classical mechanics, thermodynamics, ... was about)
-%	\item Conceptual unification vs theoretical unification
-%	\item Common standard tools (like set theory, category theory, ...)
-%	\item Connection between fields
-%	\item Reduce the number of fundamental concepts
-%	\item These more powerful tools may help theoretical unification
-%	\item realm of validity of the different theoretical theories
-%	\item realm of validity of the underlying mathematical tools
-%\end{itemize}
-
-\section{Conclusion}
-
-We have shown how the reverse physics approach, trying to find different premises from which to rederive physical theories or specific results, can help us improve our understanding.
-In the first example we showed how we can take a single theory (classical Hamiltonian mechanics) and characterize it in several ways consistent with each other (deterministic evolution that is either dynamically or thermodynamically reversible, conservation of information entropy, conservation of measurement precision, ...).
-We have seen how this helps foster a unified understanding of physics, and puts physics (and not mathematics) at the center of the discussion.
-We then showed how we can isolate a specific result of a theory (the uncertainty relationship), see what specific part of the theory is responsible for that result (fixing the entropy) and showed how another theory (classical mechanics) can be modified to achieve the same result.
-Lastly, we saw how we can recast a mostly phenomenological condition (the zero entropy of a crystalline structure at zero temperature) into a conceptual one (the zero entropy of an empty system).
-
-We believe this type of work can be beneficial in several ways.
-At the very least, it can have a significant impact within physics education.
-Most importantly, it leads to common ideas and definitions that span different fields of physics, which may help develop common tools, much like set theory and category theory provide standard tools for all of mathematics.
-Lastly, as it helps us understand the realm of validity of the different theories and also the underlying mathematical tools, it may give insight into the development of new tools and new theories.
-
-More than the particular results presented in this paper, it is the type of approach that we want to promote.
-We hope others will present their own reverse physics results, and that this may lead to a renewed interest in the manner in which physicists in the past tried to formulate the foundations of our field: through physical principles, laws and assumptions.
-
-\section*{Appendix: Proofs and calculations}
+\section*{Proofs and calculations}
 

### 2022-03-06 Articles/2020-QuantumLogic/Source/quantumlogic.tex
QuantumLogic: response to referee comments from Andrea

--- 2021-10-18 Articles/2020-QuantumLogic/Source/quantumlogic.tex
@@
 
-\author{Andrea Oldofredi\thanks{Section de Philosophie, Universit\'e de Lausanne, Switzerland.
-E-mail: Andrea.Oldofredi@unil.ch} \and Gabriele Carcassi\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109.
-E-mail: carcassi@umich.edu} \and Christine A.
-Aidala\thanks{Physics Department, University of Michigan, Ann Arbor, MI 48109.
-E-mail: caidala@umich.edu}}
+\author{The Authors}
 
@@
 \begin{abstract}
-It is generally believed that quantum mechanics entails a rejection of classical logic.
+\textcolor{blue}{At the onset of quantum mechanics, it was argued that the new theory would entail a rejection of classical logic.
 The main arguments to support this claim come from the non-commutativity of quantum observables, which allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions.
-In this essay we show that such arguments contain serious flaws.
-In particular, it is possible to keep the classical distributivity law in the context of quantum theory once the actual meaning of quantum propositions is unambiguously defined and properly taken into account.
-We moreover argue that the existence of quantum superpositions does not lead to the definition of a new rule for logical disjunctions.
-Furthermore, contrary to the received view in quantum logic, we explain that quantum mechanics can generate a distributive lattice of propositions.
-Finally, the remarkable formal similarities between the logical structures of classical and quantum mechanics are highlighted and discussed.
+While the quantum logic program is not as popular as it once was, a crucial question remains unsettled: what is the relationship between the logical structures of classical and quantum mechanics?
+In this essay we answer this question by showing that the original arguments promoting quantum logic contain serious flaws, and that quantum theory does satisfy the classical distributivity law once the full meaning of quantum propositions is properly taken into account.
+Moreover, we show that quantum mechanics can generate a distributive lattice of propositions, which, unlike the one of quantum logic, includes statements about expectation values which are of undoubtable physical interest.
+Lastly, we show that the lattice of statistical propositions in classical mechanics follows the same structure, yielding an analogue non-commutative sublattice of classical propositions.
+This fact entails that the purported difference between classical and quantum logic stems from a misconstructed parallel between the two theories.}
 \vspace{4mm}
@@
 
-As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus.\footnote{For an introduction and a clear discussion of quantum logic the reader may refer to cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{deRonde:2016}, \cite{Wilce:2021}.} This belief stood the test of time, and it is still widespread to the present day.
-For example, \cite{Giuntini:2018} claim that ``[t]he \emph{logic} of classical physical objects is naturally based on a two-valued semantics.
-Such a dichotomic situation breaks down in quantum theory'' (p.\ 4), implying that the latter entails a revision of classical propositional calculus.
-Such a view is shared by \cite{Wilce:2021}, who says: ``[m]athematically, quantum mechanics can be regarded as a non-classical probability calculus resting upon a non-classical propositional logic''.
-\noindent Similarly, presenting the standard formulation of quantum theory and its peculiar features, Stefanovich states that
-\begin{quote}
-the true logical relationships between results of measurements are different from the classical laws of Aristotle and Boole.
-The usual classical logic needs to be replaced by the so-called quantum logic (\cite{Stefanovich:2019}, p.\ 2)
-\end{quote}
-\noindent Even more explicitly, introducing the quantum logical formalism Holik writes that
-\begin{quote}
-[t]he connectives ``$\vee$'', ``$\wedge$'' and ``$\neg$'' are the quantum logical analogues of the Boolean connectives.
-It is important to remark that, with these connectives, the lattice of closed subspaces does not satisfies [\emph{sic}] the distributivity law [...] and then, it is non-Boolean.
-This implies that the new connectives cannot be interpreted in a classical way (\cite{Holik:2021}, p.\ 5).
-\end{quote}
-
-The first modifications of Classical Logic (CL) were advanced in the thirties in order to faithfully represent the physical content of quantum theory.
+As soon as Quantum Mechanics (QM) achieved a definite and coherent mathematical formulation, several physicists and philosophers claimed that its formal structure does not conform to the laws of classical propositional calculus.\footnote{For an introduction and a clear discussion of quantum logic the reader may refer to cf.\ \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{deRonde:2016}, \cite{Wilce:2021}.} The first modifications of Classical Logic (CL) were advanced in the thirties in order to faithfully represent the physical content of quantum theory.
 For instance, in 1931 the Polish philosopher Zygmunt Zawirski proposed to apply \L ukasiewicz's three-valued logic to QM, starting from considerations about the wave-particle duality and Heisenberg's uncertainty principle.
@@
 A few years after the publication of this paper, the research in quantum logic saw a notable development in both the scientific and philosophical communities, opening new research programs searching for the correct logical structures to understand and describe the physics of QM (cf.\ \cite{Reichenbach:1944}, \cite{Mackey:1957}, \cite{Finkelstein:1963}, \cite{Kochen:1965}, \cite{Jauch:1969}, \cite{Giuntini:2002}, \cite{DallaChiara:2004}, \cite{Pitowsky:2006}, \cite{Giuntini:2018}, \cite{Svozil:2020}, \cite{Holik:2021}), and intense metaphysical debates regarding the nature of logic itself and its empirical status (cf.\ notably \cite{Quine:1951}, \cite{Putnam:1968}, \cite{Dummett:1976}, \cite{Hallett:1982}, \cite{Weingartner:2004}, \cite{Bacciagaluppi:2009}).
-In addition, it is worth noting that research in quantum logic is still very prolific to the present day with many interesting applications in the fields of quantum information, computation and cryptography.
-Therefore, although QL is no longer viewed as a solution to the foundational problems affecting quantum theory\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
+
+The belief according to which QM entails a new logic stood the test of time, and it is still \textcolor{blue}{alive} to the present day.
+For example, \cite{Giuntini:2018} claim that ``[t]he \emph{logic} of classical physical objects is naturally based on a two-valued semantics.
+Such a dichotomic situation breaks down in quantum theory'' (p.\ 4), implying that the latter entails a revision of classical propositional calculus.
+Such a view is shared by \cite{Wilce:2021}, who says: ``[m]athematically, quantum mechanics can be regarded as a non-classical probability calculus resting upon a non-classical propositional logic''.
+\textcolor{blue}{Similarly, explaining the difference between the logic of classical mechanics and the logic obeyed by quantum propositions, \cite{Holik:2019} write that
+\begin{quote}
+[i]n classical and quantum mechanics, the physical properties of a system are endowed with a lattice structure. These structures are different in the classical
+and quantum case, and they determine the logical structure of the physical system (\cite{Holik:2019}, p.\ 363-364).
+\end{quote}
+\noindent More precisely the authors say:
+\begin{quote}
+The distributive inequalities are the main difference between classical and quantum logic.
+In the classical lattice, all properties satisfy the distributive equalities, but in the quantum lattice, only distributive inequalities hold, in general (\emph{ibid.}).
+\end{quote}
+}
+
+\textcolor{blue}{Even more explicitly, presenting the standard formulation of quantum theory and its peculiar features, Stefanovich states that
+\begin{quote}
+the true logical relationships between results of measurements are different from the classical laws of Aristotle and Boole.
+The usual classical logic needs to be replaced by the so-called quantum logic (\cite{Stefanovich:2019}, p.\ 2).
+\end{quote}
+}
+
+ \textcolor{blue}{While the quantum logic programme is still being actively pursued by a sizable community, and despite the many interesting applications in the fields of quantum information, computation and cryptography that are being investigated to the present day, QL is no longer viewed as a solution to the foundational problems affecting quantum theory.\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
 Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory].
-This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).}, it remains an important discipline of contemporary physics.
-In this paper we will neither take a stance in the dispute concerning the ontological status of logic and its laws, nor we will consider recent developments of QL given that they are not strictly relevant to the point we are going to make.
-Indeed, the present essay is focused on the logical structures common to both quantum and classical mechanics, following the spirit of the former discussion.\footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM.
-First-order and higher-order logics will not be discussed here, nor the logic of relativistic quantum theories.}
+This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).} Nevertheless, although it may be reasonable for a physicist to simply discard the approach over practical considerations, such as the lack of a desired result, on philosophical grounds this move would beg important questions, as for instance: what is the relationship between the logical structures of classical and quantum mechanics?
+How are they different and how are they similar?
+What should we think about the core insights that led to the development of QL?
+Were they substantially correct, or should they be rejected?
+We think that providing a correct answer to these issues will lead (i) to a deeper understanding of the structural relations existing between classical logic and quantum mechanics---which unfortunately are not yet precisely understood---and (ii) to a clarification of important confusions concerning the motivations behind QL.
+%For these reasons, the present essay is focused on the logical structures common to both quantum and classical mechanics, highlighting the similarities between these two theoretical frameworks.
+ \footnote{In the present essay we will be concerned only with propositional calculus in the context of non-relativistic QM.
+First-order and higher-order logics will not be discussed here, nor the logic of relativistic quantum theories.}}
 
-Referring to this, one of us recently faced the question of whether quantum physics implies a revision of classical logic (cf.\ \cite{Oldofredi:2020}).
+Referring to this, one of us recently faced the question of whether quantum physics implies a revision of classical logic (cf.\ [Reference Deleted]).
 Alternatively stated, it was asked if classical logic must be inevitably abandoned in the quantum realm, or if there are quantum theories compatible with classical propositional calculus.
@@
 
-% In the following section we will see that Hilbert spaces already posses a mathematical structure that follows classical logic, namely the Borel $\sigma$-algebra, and therefore quantum mechanics already posses a propositional lattice that follows classical logic.
-Moreover, we show that classical probabilistic theories also feature a non-distributive lattice that is the analogue of quantum logic, which means non-distributivity is neither required nor unique to quantum mechanics.
-
 \section{Quantum Disjunction and Distributivity Law}
@@
 In this section we will review the examples given in the previous one and explain that the failure of distributivity stems from the ambiguity of the quantum logical propositions and, once that is clarified, standard classical rules can be applied also in the quantum context.
-More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members is, and (ii) that \eqref{DL} can be retained in QM, since the arguments seen above---usually meant to show its failure---are flawed.
+\textcolor{blue}{More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction can be true when neither of its members is\footnote{\textcolor{blue}{Clearly, a quantum disjunction is false when both disjuncts are false; however, the interesting case for our discussion is the one in which quantum propositions have undetermined truth values.}}---i.e.\ when both disjuncts have undetermined truth values as in the case discussed above in which we considered a system in a linear combination of $y$-spin states}---and (ii) that \eqref{DL} can be retained in QM, since the arguments seen above---usually meant to show its failure---are flawed.
 
@@
 \end{description}
-\noindent Because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true.
-Therefore $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$.
-Consequently $q_i \vee r_i = q'_i \vee r'_i$.
+\noindent \textcolor{blue}{If $\rho_i$ is the (potentially) mixed state before the measurement, the expectation value for $y$-spin is given by $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar \langle \psi_y^+ | \rho_i | \psi_y^+ \rangle - \frac{1}{2} \hbar \langle \psi_y^- | \rho_i | \psi_y^- \rangle$.
+Therefore $q'_i$ is true if and only if $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar$.
+Because $\rho_i$ is positive definite, this can only happen if $\langle \psi_y^+ | \rho_i | \psi_y^+ \rangle = 1$ and $\langle \psi_y^- | \rho_i | \psi_y^- \rangle = 0$, which means $\rho_i = | \psi_y^+ \rangle \langle \psi_y^+ |$.
+This entails that the expectation value for $y$-spin is $\frac{1}{2} \hbar$ if and only if the electron was prepared in a pure state of $y$-spin up.
+In other words, because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true.
+We have that $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$.} Consequently $q_i \vee r_i = q'_i \vee r'_i$.
 Note that $q'_i \vee r'_i$ is the proposition ``before the measurement, the expectation value for $y$-spin is $\pm \frac{1}{2} \hbar$''.
@@
 Therefore it is not the case that the quantum disjunction can be true even if none of the elements is true.
+\textcolor{blue}{The overall point is that the incompatibility of the observables does indeed introduce a problem because it is not clear how to evaluate propositions that mix different variables.
+The use of equivalent propositions over expectation values helps us solve the ambiguities.}
 
@@
 
-Let us now turn our attention to distributivity and try to rewrite \eqref{contradiction} by using the refined propositions that explicitly include the temporal dependence.
+\textcolor{blue}{Several authors have already argued in different ways that it is possible to retain distributivity in quantum mechanics.
+For instance, \cite{Park:1968} disputed the identification between Hermitian operators and quantum observables and provided a new interpretation of quantum measurement theory rejecting the notion of incompatible observables---and thereby incompatible measurements---which, as we have seen, constitutes the basis to show the failure of the distributivity law in the quantum realm.
+Remarkably, also Bohr argued that classical logic should be maintained in quantum theory as a consequence of the complementarity principle:
+\begin{quote}
+The aim of the idea of complementarity was to allow of keeping the usual logical forms while procuring the extension necessary for including the new situation relative to the problem of observation in atomic physics (Bohr quoted in \cite{Faye:2021}, p.\ 115).
+\end{quote}
+Another interesting quote reported in \cite{Faye:2021} clearly shows that the Danish physicists strongly opposed the idea of replacing classical propositional calculus with QL: 
+\begin{quote}
+The question has been raised whether recourse to multivalued logics is needed for a more appropriate representation of the situation.
+From the preceding argumentation it will appear, however, that all departures from common language and ordinary logic are entirely avoided by reserving the word `phenomenon' solely for reference to unambiguously communicable information, in the account of which the word `measurement' is used in its plain meaning of standardized comparison (Bohr quoted in \cite{Faye:2021}, p.\ 115).
+\end{quote}
+Although Bohr did not provide formal arguments against the introduction of quantum logic, his views express the idea according to which macroscopic measurement results must be described with ordinary language, therefore, they must be subjected to the rules of classical logic---in his view the intrinsic novelties of quantum mechanics have to be found in the contextual and complementary nature of quantum phenomena on the one hand, and the entanglement between quantum systems and macroscopic devices which in measurement situations form an indissoluble unity on the other.
+Hence, given that quantum propositions concern only measurement results, Bohr concluded that they should be governed by the laws of classical propositional calculus.
+To this regard, \cite{Heelan:1970} provided a formalization of Bohr's arguments based on the complementarity principle and showed that the statements about a single quantum mechanical event, i.e.\ a measurement of a certain observable given a precise experimental context, do follow a classical propositional calculus.}\footnote{\textcolor{blue}{We thank one anonymous reviewer for having pointed out Heelan's work to us.}} \textcolor{blue}{More precisely, Heelan claims that QM introduces a distinction between events, corresponding to the actual performance of a certain measurement, and physical contexts, i.e.\ experimental settings determining the necessary and sufficient conditions for the realization of a particular measurement outcome.
+In his view, then, one has to introduce two different languages: an event-language useful to formally describe a particular observation of a ccertain quantum observable and a context-language, which is a meta-language necessary in order to properly speak about how a certain event can possibly occur.\footnote{\textcolor{blue}{Heelan exemplifies such a distinction as follows: ``If the event, for example, is a particle-location event, the event-language is position-language, and the physical context is a standardized instrumental set-up plus whatever other physical conditions are necessary and sufficient for the measurement of a given range of possible particle-position events'' (\cite{Heelan:1970}, p.\ 96).}} If we consider only propositions involving and relating incompatible contexts, Heelan says, then we will generate a lattice of propositions which is not distributive, however, if we take into account single event-languages---each of which is coupled with a given context-language---the logic of such individual quantum events can be classical.
+This captures Bohr's idea according to which the principle of complementarity arises from the context-dependent character of quantum mechanical events.}
+
+\textcolor{blue}{Although we partially agree with Heelan's view in saying that propositions about individual quantum measurements can generate a distributive lattice, we want to generalize such a claim.
+In fact, we show---with simpler arguments and independently of Bohr's interpretation of quantum theory---that taking into account the temporal order of individual quantum measurements one can relate even statements about incompatible observations into a distributive lattice.
+In addition, we show that ignoring the temporal order of measurements may lead to violations of the distributivity law also in classical mechanics; thus, it is our aim to argue that once the correct interpretation of quantum propositions is taken into account the lattice generated by such statements is distributive.}
+
+In order to derive our desired result, let us now turn our attention to distributivity and try to rewrite \eqref{contradiction} by using the refined propositions that explicitly include the temporal dependence.
 The proposition $p$ always referred to preparation, therefore it maps to $p_i$ in all cases.
@@
 
-\begin{figure}
-
 \begin{center}
 \includegraphics[scale=.8]{Ball.eps}
+\begin{quote}
+\small{Figure 2: Graphic representation of the discussed example.
+Initially a ball is in position $p$ over a hatch.
+After the hatch opens, the ball drops and falls to either position $q$ or $r$ with equal chance.}
+\end{quote}
 \end{center}
-\caption{\small{Graphic representation of the discussed example.
-Initially a ball is in position $p$ over a hatch.
-After the hatch opens, the ball drops and falls to either position $q$ or $r$ with equal chance.}}
-\end{figure}
 
@@
 Therefore $V$ is not a closed subspace.
-Our proposition, despite being physically interesting, does not correspond to a closed subspace but does correspond to a Borel set.
+Our proposition, despite being physically interesting, does not correspond to a closed subspace but does correspond to a Borel set.\footnote{\textcolor{blue}{Note that we are not claiming that all Borel sets can be constructed in this way or that all Borel sets correspond to physically interesting propositions.
+First of all, it is not true: $\{ \psi \}$ and $\{ e^{\imath \theta} \psi\}$ are different Borel sets, yet they are not physically distinguishable since a difference in absolute phase is not physically relevant.
+This can be fixed by considering the Borel sets of the projective space.
+Second, to really examine the physicality of all Borel sets we need a ``general theory of experimental logic'', of the type provided by \cite{Kelly:1996} or \cite{Carcassi:2021}, which would go well beyond the present discussion.
+With appropriate caveats (e.g.
+the Hilbert space must be separable), one can say that the Borel sets correspond to propositions that can be associated to an experimental test (what \cite{Carcassi:2021} call ``theoretical statements'') which may or may not terminate in any or all cases.
+We leave this discussion for another work.}}
 

### 2022-03-08 Articles/2020-QuantumLogic/Source/response1.tex
QuantumLogic: final version of the response

@@
-\documentclass[11pt]{article}
-
-\usepackage[margin=1.25in]{geometry}
-
-\usepackage{amsmath, amsthm, amssymb}
+\documentclass[11pt, executivepaper]{article}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{natbib}
+\usepackage{amsmath}
+\usepackage{xcolor}
 \usepackage{amsfonts}
-\usepackage{MnSymbol}
-\usepackage{mathtools}
-\usepackage{xcolor}
 \usepackage{graphicx}
 \usepackage{enumitem}
-
-\newcommand\change[1]{{\color{red!75!black} #1}}
-
+\usepackage{geometry}
+ \geometry{
+ a4paper,
+ total={158mm,237mm},
+ left=28mm,
+ top=28mm,
+ }
+\usepackage{hyperref}
+\hypersetup{colorlinks= true, allcolors=blue}
+\setcitestyle{aysep={}}
 \begin{document}
 
-TO THE EDITOR
+\title{\textbf{Replies to Reviewers' Comments}}
 
-thank you very much for having provided us
-good quality referee reports for our paper. We appreciate that this is
-not an easy task.
+\author{The Authors}
 
-\section*{Referee 1}
+\maketitle
 
+\noindent We would like to thank the Editors for the opportunity to revise the manuscript; moreover, we also warmly thank both reviewers for their positive and helpful comments which led to significant changes and improvements of our paper.
+We appreciate that, we know that this is not an easy task.
+\vspace{2mm}
+
+In order to make the revisions immediately visible, every change in the manuscript has been written in blue so that it is highlighted with respect to the text. 
+\vspace{2mm}
+
+Finally, we hope to have properly addressed the doubts and remarks of both referees with these answers and the related changes made in the manuscript.
+
+\section{Replies to Reviewer 1}
+
+\begin{enumerate}
+\item \textbf{Reviewer 1:} p.
+2: The quotation from Holik seems quite reasonable.
+All that Holik claims is that the connectives he is using in the context of quantum theory are ``analogues'' to the classical truth-functional connectives, and immediately mentions that they cannot be taken to be the classical connectives.
+Unless Holik adds that the quantum ones are meant to replace or correct classical logic, the statement seems fine.
+It is certainly much better than the Stefanovich talk of ``replacement''.
+\vspace{2mm}
+
+\textbf{Authors:} We agree with Reviewer 1 that this quote is not very representative of the claim that QM requires a new logic, or that QL is still alive to the present day.
+In order to provide more relevant textual evidence, we decided to remove Holik's claim in favor of a passage taken from \cite{Holik:2019}.
+In this essay the authors argue (i) that quantum mechanics does require a logic that is non-classical in virtue of the mathematical formalism underlying the theory, and (ii) that it is possible to make a transition from QL to classical propositional calculus taking the classical limit of QM.
+We think that this essay represents very clearly the idea expressed in the introduction of our paper.
+
+The new text is the following:
+
+\textcolor{blue}{Similarly, explaining the difference between the logic of classical mechanics and the logic obeyed by quantum propositions, \cite{Holik:2019} write that
 \begin{quote}
-The topic of this paper is the standing of "quantum logic" vis-á-vis "classical logic", and in particular whether quantum mechanics demands either a rejection of or adjustment to classical logic.
-The answer, in a nutshell, is "no".
-The burden of the paper is to explain exactly why the standard arguments to the opposite conclusion are incorrect, and incidentally how both distributive Boolean lattices of propositions can be defined in quantum mechanics and non-distributive, non-Boolean lattices of propositions can be defined in the theory of probability distributions over classical phase space.
-These observations serve to show how the non-distributivity is not connected either to quantum mechanics or to the use of complex (rather than real) numbers.
+[i]n classical and quantum mechanics, the physical properties of a system are endowed with a lattice structure. These structures are different in the classical
+and quantum case, and they determine the logical structure of the physical system (\cite{Holik:2019}, p.\ 363-364).
+\end{quote}
+\noindent More precisely the authors say:
+\begin{quote}
+The distributive inequalities are the main difference between classical and quantum logic.
+In the classical lattice, all properties satisfy the distributive equalities, but in the quantum lattice, only distributive inequalities hold, in general (\emph{ibid.}).
+\end{quote}
+}
 
-The paper is clear and decisive.
-There are sufficiently many citations to justify publications, for even though the star of quantum logic has faded it has not entirely gone out.
-It is worthwhile to have all of this laid out in a single, compact presentation.
+\item \textbf{Reviewer 1:} p.
+9: ``More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction must be true when neither of its members is''.
+But surely that is attacking a straw man.
+No one every claimed that a ``quantum disjunction'' (join) must be true when the propositions joined fail to be true!
+For example, when they are both false.
+\vspace{2mm}
 
-I have only a couple of comments/ suggestions.
+\textbf{Authors:} We agree that our statement at page 9 may seem ambiguous, and we have changed it in the revised version of the essay.
+The original statement was referring to the quote of \cite{deRonde:2016} at page 6 of the first submission.
+In the revised version we write: ``\textcolor{blue}{More precisely, we are going to argue that (i) it is not necessarily the case that a quantum disjunction can be true when neither of its members is---i.e.\ when both disjuncts have undetermined truth values as in the case discussed above in which we considered a system in a linear combination of $y$-spin states}''.
+Furthermore, we added the following footnote: ``\textcolor{blue}{Clearly, a quantum disjunction is false when both disjuncts are false; however, the interesting case for our discussion is the one in which quantum propositions have undetermined truth values}''.
+In this manner, we hope to have eliminated the straw man.
 
-p.
-2: The quotation from Holik seems quite reasonable.
-All that Holik claims is that the connectives he is using in the context of quantum theory are "analogues" to the classical truth-functional connectives, and immediately mentions that they cannot be taken to be the classical connectives.
-Unless Holik adds that the quantum ones are meant to replace or correct classical logic, the statement seems fine.
-It is certainly much better than the Stefanovich talk of "replacement".
+\item \textbf{Reviewer 1:} p. 11: ``The solution is simple: while the value of an observable is not always defined, its expectation value always is. The idea is
+then to translate $q_i$ and $r_i$ into propositions about expectation values, which we already know how to handle, and apply the standard rules. Consider the following statements:
+\begin{itemize}
+\item $q'_i$: before the measurement, the expectation value for y-spin is $\frac{1}{2} \hbar$.
 
-p. 9: "More precisely, we are going to argue that (i) it is not necessarily the case that a quantum
-disjunction must be true when neither of its members is".
-But surely that is attacking a straw man.
-No one every claimed that a "quantum disjunction" (join) must be true when the propositions joined fail to be true!
-For example, when they are both false.
-
-\end{quote}
-
-TODO
-
-\begin{quote}
-p. 1 1: "The solution is simple: while
-the value of an observable is not always defined, its expectation value always is. The idea is
-then to translate qi and ri into propositions about expectation values, which we already know
-how to handle, and apply the standard rules. Consider the following statements:
-
-q'i: "before the measurement, the expectation value for y-spin is ½ h.
-
-r'i: "before the measurement, the expectation value for y-spin is -½ h.
-
-Because the spin of the electron along any axis is bounded by ±1/2 h, the only way that q'I is
-true is if qi is true. Therefore q'I and r'i are respectively equivalent to qi and ri."
+\item $r'_i$: before the measurement, the expectation value for y-spin is $-\frac{1}{2} \hbar$.
+\end{itemize}
+Because the spin of the electron along any axis is bounded by $\pm\frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true.
+Therefore $q'_i$ and $r'_i$ are respectively equivalent to $q_i$ and $r_i$.''
 
 I can't follow this claim.
-It is true that the only way that the expectation value can be, say, ½ h is if the system is in an eigenstate of y-spin with the value up, but one could certainly imagine a situation (according to a theory) in which the particle does indeed have y-spin up but the expectation value associated with the system is not +1/2 h.
+It is true that the only way that the expectation value can be, say, $\frac{1}{2} \hbar$ is if the system is in an eigenstate of y-spin with the value up, but one could certainly imagine a situation (according to a theory) in which the particle does indeed have y-spin up but the expectation value associated with the system is not $+1/2 h$.
 This could be true on a statistical understanding of the expectation value: even though this system has y-spin up, the expectation value is calculated with respect to an ensemble of systems, not all of which have y-spin up.
 So the inference certainly goes through in one direction, but whether it does in the other depends on how the expectation values are understood.
-But "equivalent" suggests equivalence in both directions.
+But ``equivalent'' suggests equivalence in both directions.
 So I think more would have to be said here.
-\end{quote}
 
-To remove ambiguity, we expanded the example to take the most general case, that of a density matrix, and show that fixing the expectation value to 1/2 h means the density matrix must correspond to the pure state of spin up.
-That way it is clear what we mean by expectation, and we cover all possible cases that are present in quantum mechanics (i.e.
-there is no mixed state that could give the same result).
-It reads:
+\vspace{2mm}
 
-\change{If $\rho_i$ is the (potentially) mixed state before the measurement, the expectation value for $y$-spin is given by $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar \langle \psi_y^+ | \rho_i | \psi_y^+ \rangle - \frac{1}{2} \hbar \langle \psi_y^- | \rho_i | \psi_y^- \rangle$.
+\textbf{Authors:} To remove ambiguity, we expanded the example to take the most general case, that of a density matrix, and show that fixing the expectation value to $\frac{1}{2} \hbar$ entails that the density matrix must correspond to the pure state of spin up.
+In this manner it should be clearer what we mean by expectation, and we cover all possible cases that are present in quantum mechanics (i.e.\ there is no mixed state that could give the same result).
+We added the following paragraph:
+
+\textcolor{blue}{If $\rho_i$ is the (potentially) mixed state before the measurement, the expectation value for $y$-spin is given by $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar \langle \psi_y^+ | \rho_i | \psi_y^+ \rangle - \frac{1}{2} \hbar \langle \psi_y^- | \rho_i | \psi_y^- \rangle$.
 Therefore $q'_i$ is true if and only if $\text{Tr}(\rho_i S_y) = \frac{1}{2} \hbar$.
 Because $\rho_i$ is positive definite, this can only happen if $\langle \psi_y^+ | \rho_i | \psi_y^+ \rangle = 1$ and $\langle \psi_y^- | \rho_i | \psi_y^- \rangle = 0$, which means $\rho_i = | \psi_y^+ \rangle \langle \psi_y^+ |$.
-This means that the expectation value for $y$-spin is $\frac{1}{2} \hbar$ if and only if the electron was prepared in a pure state of $y$-spin up.
+This entails that the expectation value for $y$-spin is $\frac{1}{2} \hbar$ if and only if the electron was prepared in a pure state of $y$-spin up.
 In other words, because the spin of the electron along any axis is bounded by $\pm \frac{1}{2} \hbar$, the only way that $q'_i$ is true is if $q_i$ is true.
@@
 
-Note that the argument holds regardless of how one interprets $\rho$ (e.g. state of knowledge, fraction of repeated experiments, ...).
+Note that the argument holds regardless of how one interprets $\rho$ (e.g.\ state of knowledge, fraction of repeated experiments, ...). 
+\vspace{2mm}
+
+Reviewer 1 then adds:
 
@@
 \end{quote}
+
 Actually, the smoother transition is a bonus.
-The main reason to introduce expectation values here is to settle the question of how to evaluate statements of over one observable when an incompatible observable is prepared.
-As the text reads ``The difficulty here is that before the measurement of $S_y$ only $x$-spin is defined, so how are we going to compute statements about the value of $y$-spin?'' We suspect that, given the length of the paragraph, the point is lost by the end, so we close the paragraph with:
+The main reason to introduce expectation values here is to settle the question of how to evaluate statements about a certain observable when another operator incompatible with it is prepared.
+As the text reads ``The difficulty here is that before the measurement of $S_y$ only $x$-spin is defined, so how are we going to compute statements about the value of $y$-spin?'' We suspect that, given the length of the paragraph, the point is lost by the end, so we close the paragraph with a new proposition:
 
-\change{The overall point is that the incompatibility of the observables does indeed introduce a problem because it is not clear how to evaluate propositions that mix different variables.
+\textcolor{blue}{The overall point is that the incompatibility of the observables does indeed introduce a problem because it is not clear how to evaluate propositions that mix different variables.
 The use of equivalent propositions over expectation values helps us solve the ambiguities.}
 
-So the idea is that it is useful to add these propositions anyway, so it already gives us a hint to compare to classical statistical propositions.
+Our idea is that it is useful to add these propositions anyway, since it already gives us a hint to compare quantum propositions to classical statistical propositions.
+\end{enumerate}
 
-\section*{Referee 2}
+We hope with these answers to have properly replied to the useful comments we received from reviewer 1.
 
-\begin{quote}
-This is a very interesting article, and it puts back on the table an older issue that had been dormant for a while.
-As such, it is a helpful contribution.
-However, it does have several shortcomings that prevent me from endorsing for publication in its present form.
-\end{quote}
+\section{Replies to Reviewer 2}
 
-\begin{quote}
-1.
-The primary defect of the paper is that it greatly overstates the degree to which people are convinced that QM demands a revision of our logic.
-I would say that 90\% of philosophers of physics believe that it does *not* demand a revision of our logic, and that 90\% of physicists would say that they do not know what it means to say that QM demands a revision of logic.
+\begin{enumerate}
+\item \textbf{Reviewer 2:} The primary defect of the paper is that it greatly overstates the degree to which people are convinced that QM demands a revision of our logic.
+I would say that 90\% of philosophers of physics believe that it does \textbf{not} demand a revision of our logic, and that 90\% of physicists would say that they do not know what it means to say that QM demands a revision of logic.
 The citations that the authors use represent only a small minority of philosophers and physicists who were trying to sell this idea.
@@
 The main problem is that their arguments are largely aimed at a straw person -- and hence it is not so likely that the paper would receive a broad readership.
-\end{quote}
+\vspace{2mm}
 
-Looking again at the written prose, the stated motivation of the paper (particularly the abstract) is indeed just an argument against Quantum Logic.
+\textbf{Authors:} Looking again at the written prose, the stated motivation of the paper (particularly the abstract) is indeed just an argument against Quantum Logic.
 In that sense, the referee is correct.
-On one hand, QL still attracts funding and interest and, as referee 1 puts it:
+On one hand, QL still attracts funding and interest and, as reviewer 1 puts it:
 \begin{quote}
@@
 \end{quote}
-So we still think talking about QL is relevant, However, as the referee 2 notes later, the article goes past QL and compares at a deeper level the logical structures of classical and quantum mechanics.
+So we still think talking about QL is relevant (cf.\ also the reply to the first comment of reviewer 1).
+However, as reviewer 2 notes later, the article goes past QL and compares at a deeper level the logical structures of classical and quantum mechanics.
 This, we believe, is of general interest regardless of how one may feel about QL.
-Given  comments from both referees, we propose to keep the arguments against QL, but to make it clear that our discussion goes deeper.
+Given the comments from both referees, we propose to keep the arguments against QL, but to make it clear that our discussion goes deeper.
 
-We revised the abstract as follows:
+In turn, we revised the abstract as follows:
 
-\change{At the onset of quantum mechanics, it was speculated that the new theory would entail a rejection of classical logic.
+\textcolor{blue}{At the onset of quantum mechanics, it was argued that the new theory would entail a rejection of classical logic.
 The main arguments to support this claim come from the non-commutativity of quantum observables, which allegedly would generate a non-distributive lattice of propositions, and from quantum superpositions, which would entail new rules for quantum disjunctions.
-While the quantum logic program is not as popular as it once was, the core question remains unsettled: what is the relationship between the logical structures of the two theories?
-In this essay we show that the original arguments promoting quantum logic contain serious flaws, and that quantum theory does satisfy the classical distributivity law once the full meaning of quantum propositions is properly taken into account.
-Moreover, we show that quantum mechanics can generate a distributive lattice of proposition, which, unlike the one of quantum logic, includes statements about expectation values which are of undoubtable physical interest.
+While the quantum logic program is not as popular as it once was, a crucial question remains unsettled: what is the relationship between the logical structures of classical and quantum mechanics?
+In this essay we answer this question by showing that the original arguments promoting quantum logic contain serious flaws, and that quantum theory does satisfy the classical distributivity law once the full meaning of quantum propositions is properly taken into account.
+Moreover, we show that quantum mechanics can generate a distributive lattice of propositions, which, unlike the one of quantum logic, includes statements about expectation values which are of undoubtable physical interest.
 Lastly, we show that the lattice of statistical propositions in classical mechanics follows the same structure, yielding an analogue non-commutative sublattice of classical propositions.
-This shows that the purported difference between classical and quantum logic stems from a misconstructed parallel between the two theories.}
+This fact entails that the purported difference between classical and quantum logic stems from a misconstructed parallel between the two theories.}
 
-We revised part of the introduction as follows:
+In addition, we revised part of the introduction as follows:
 
-\change{While the approach is still being actively pursued by a sized community, and despite the many interesting applications in the fields of quantum information, computation and cryptography that are being investigated to the present day, QL is no longer viewed as a solution to the foundational problems affecting quantum theory.\footnote{This conclusion is now accepted among experts; for details see e.g.\ \cite{Bacciagaluppi:2009} and \cite{Giuntini:2002}.
-Referring to this, the latter authors stated explicitly that ``quantum logics are not to be regarded as a kind of ``clue'', capable of solving the main physical and epistemological difficulties of QT [quantum theory].
-This was perhaps an illusion of some pioneering workers in quantum logic'' (\cite{Giuntini:2002}, p.\ 225).} Nevertheless, while it may be reasonable for a physicist to simply discard the approach over practical considerations, such as the lack of a desired result, on philosophical grounds it begs the question: what is the relationship between the logical structures of classical and quantum mechanics?
+\textcolor{blue}{While the quantum logic programme is still being actively pursued by a sizable community, and despite the many interesting applications in the fields of quantum information, computation and cryptography that are being investigated to the present day, QL is no longer viewed as a solution to the foundational problems affecting quantum theory.
+Nevertheless, although it may be reasonable for a physicist to simply discard the approach over practical considerations, such as the lack of a desired result, on philosophical grounds this move would beg important questions, as for instance: what is the relationship between the logical structures of classical and quantum mechanics?
 How are they different and how are they similar?
-What should we think about the core insights that lead to the development of QL?
+What should we think about the core insights that led to the development of QL?
 Were they substantially correct, or should they be rejected?
-Given the fundamental role that logic plays in philosophy, we feel that these questions still must be answered regardless of how one may feel towards QL, or we are bound to make the same mistakes.}
+We think that providing a correct answer to these issues will lead (i) to a deeper understanding of the structural relations existing between classical logic and quantum mechanics---which unfortunately are not yet precisely understood---and (ii) to a clarification of important confusions concerning the motivations behind QL.}
 
-TODO: Maybe review the conclusions as well?
-
-\begin{quote}
-2.
-In Section 3, the authors argue the logic of quantum propositions is distributive ...
+\item \textbf{Reviewer 2:} In Section 3, the authors argue the logic of quantum propositions is distributive ...
 if one indexes the propositions so that they refer to times (either before or after measurement).
@@
 The point of all these suggestions is roughly that if one indexes propositions by their context, then a non-distributive logic can be represented as a contextual distributive logic.
+\vspace{2mm}
+
+\textbf{Authors:} In the revised manuscript we acknowledge that several authors argued that distributivity can be retained in QM.
+More precisely, we mention Bohr's ideas according to which classical logic should be maintained in the quantum realm in virtue of the complementarity principle.
+Furthermore, we explicitly mention Heelan's work on QL.
+However, our argument can be seen as a generalization of Heelan's claims and our approach is technically much simpler w.r.t.\ his.
+Hence, just for this, we believe that our argument can still be considered a positive contribution to the literature.
+In addition, while Heelan's paper heavily relies on Bohr's interpretation of QM---currently endorsed by a minority of experts in quantum foundations---our argument is independent of any interpretation of quantum theory.
+Finally, in this section we show that if one does not take into consideration the temporal order of measurements, the distributivity law is violated also in classical mechanics; to our knowledge this point has not yet been raised in the literature.
+
+In order to address this comment we added the following paragraphs to Section 3.2:
+
+\textcolor{blue}{Several authors have already argued in different ways that it is possible to retain distributivity in quantum mechanics.
+For instance, \cite{Park:1968} disputed the identification between Hermitian operators and quantum observables and provided a new interpretation of quantum measurement theory rejecting the notion of incompatible observables---and thereby incompatible measurements---which, as we have seen, constitutes the basis to show the failure of the distributivity law in the quantum realm.
+Remarkably, also Bohr argued that classical logic should be maintained in quantum theory as a consequence of the complementarity principle:
+\begin{quote}
+The aim of the idea of complementarity was to allow of keeping the usual logical forms while procuring the extension necessary for including the new situation relative to the problem of observation in atomic physics (Bohr quoted in \cite{Faye:2021}, p.\ 115).
 \end{quote}
+Another interesting quote reported in \cite{Faye:2021} clearly shows that the Danish physicists strongly opposed the idea of replacing classical propositional calculus with QL: 
+\begin{quote}
+The question has been raised whether recourse to multivalued logics is needed for a more appropriate representation of the situation.
+From the preceding argumentation it will appear, however, that all departures from common language and ordinary logic are entirely avoided by reserving the word `phenomenon' solely for reference to unambiguously communicable information, in the account of which the word `measurement' is used in its plain meaning of standardized comparison (Bohr quoted in \cite{Faye:2021}, p.\ 115).
+\end{quote}
+Although Bohr did not provide formal arguments against the introduction of quantum logic, his views express the idea according to which macroscopic measurement results must be described with ordinary language, therefore, they must be subjected to the rules of classical logic---in his view the intrinsic novelties of quantum mechanics have to be found in the contextual and complementary nature of quantum phenomena on the one hand, and the entanglement between quantum systems and macroscopic devices which in measurement situations form an indissoluble unity on the other.
+Hence, given that quantum propositions concern only measurement results, Bohr concluded that they should be governed by the laws of classical propositional calculus.
+To this regard, \cite{Heelan:1970} provided a formalization of Bohr's arguments based on the complementarity principle and showed that the statements about a single quantum mechanical event, i.e.\ a measurement of a certain observable given a precise experimental context, do follow a classical propositional calculus.}
 
-\begin{quote}
-3.
-Section 4 is actually innovative, and could well be developed into a novel point.
+\textcolor{blue}{More precisely, Heelan claims that QM introduces a distinction between events, corresponding to the actual performance of a certain measurement, and physical contexts, i.e.\ experimental settings determining the necessary and sufficient conditions for the realization of a particular measurement outcome.
+In his view, then, one has to introduce two different languages: an event-language useful to formally describe a particular observation of a ccertain quantum observable and a context-language, which is a meta-language necessary in order to properly speak about how a certain event can possibly occur.[footnote: Heelan exemplifies such a distinction as follows: ``If the event, for example, is a particle-location event, the event-language is position-language, and the physical context is a standardized instrumental set-up plus whatever other physical conditions are necessary and sufficient for the measurement of a given range of possible particle-position events'' (\cite{Heelan:1970}, p.\ 96).] If we consider only propositions involving and relating incompatible contexts, Heelan says, then we will generate a lattice of propositions which is not distributive, however, if we take into account single event-languages---each of which is coupled with a given context-language---the logic of such individual quantum events can be classical.
+This captures Bohr's idea according to which the principle of complementarity arises from the context-dependent character of quantum mechanical events.}
+
+\textcolor{blue}{Although we partially agree with Heelan's view in saying that propositions about individual quantum measurements can generate a distributive lattice, we want to generalize such a claim.
+In fact, we show---with simpler arguments and independently of Bohr's interpretation of quantum theory---that taking into account the temporal order of individual quantum measurements one can relate even statements about incompatible observations into a distributive lattice.
+In addition, we show that ignoring the temporal order of measurements may lead to violations of the distributivity law also in classical mechanics; thus, it is our aim to argue that once the correct interpretation of quantum propositions is taken into account the lattice generated by such statements is distributive.}
+
+\item \textbf{Reviewer 2:} Section 4 is actually innovative, and could well be developed into a novel point.
 Especially interesting is the argument (pages 16-18) that all Borel subsets of Hilbert space represent physically interesting propositions about the system.
 But let me note a few qualifications.
-(a) The authors do not show that *every* Borel subset of H is of the form described in Proposition 2.
+(a) The authors do not show that \textbf{every} Borel subset of $\mathcal{H}$ is of the form described in Proposition 2.
 That would be a very interesting result.
-(b) The authors do not cite others who have argued that all Borel subsets of H are physically relevant.
+(b) The authors do not cite others who have argued that all Borel subsets of $\mathcal{H}$ are physically relevant.
 See especially David Wallace's book on the quantum multiverse.
 (c) The authors do not consider any of the many objections that could be raised against their proposal.
-For example, one would not normally say that "quantity A having expectation value r" is a property of a system, but that it is a meta-property representing our ignorance of the actual value of A.
-\end{quote}
+For example, one would not normally say that ``quantity A having expectation value r'' is a property of a system, but that it is a meta-property representing our ignorance of the actual value of A.
+\vspace{2mm}
 
-For (a), the situation is a bit more complicated. Let us mention a few problems.
-
-Not all Borel subsets of H are of the form described in Proposition 2.
-In fact, they are not all physically relevant.
-The issue is that a state is not a vector in the Hilbert space, but a ray (i.e.
-a subspace of dimension 1).
+\textbf{Authors:} For (a), the situation is a bit more complicated.
+Let us mention a few problems.
+Not all Borel subsets of $\mathcal{H}$ are of the form described in Proposition 2.
+In fact, they are \emph{not} all physically relevant.
+The issue is that a state is not a vector in the Hilbert space, but a ray (i.e.\ a subspace of dimension 1).
 The absolute phase and absolute norm are not physically observable per se, though these can be differentiated with Borel sets.
@@
 Additionally: not all rays can realistically be considered physical.
-Some states, for example, correspond to infinite expectation for position, momentum, energy.
-One can, however, show that the set of states with finite expectation values for all polynomial of position and momentum corresponds to the Schwartz space, which is dense in the Hilbert space.
+Some states, for example, correspond to infinite expectations for position, momentum, energy.
+One can, however, show that the set of states with finite expectation values for all polynomials of position and momentum corresponds to the Schwartz space, which is dense in the Hilbert space.
 One can then show that a basis for the topology of the projective space of the Schwartz space can be constructed with sets of the form described in Proposition 2.
 
-More in general, to make a solid map between statements and experimental verification, the best way would be to use the results of Kelly or Carcassi/Aidala that link the open set of a topology with experimentally verifiable.
-This would also allow to go in further details.
+More in general, to make a solid map between statements and experimental verification, the best way would be to use the results of Kelly or Carcassi/Aidala that link the open set of a topology with experimentally verifiable propositions.
+This would also allow one to go in further details.
 For example, for the space to be (at least in the limit) explorable experimentally the topology must be second countable and $T_0$.
-Outside of those restriction, one can show that the points of the space do not correspond to cases that can be experimentally distinguished.
+Outside of those restrictions, one can show that the points of the space do not correspond to cases that can be experimentally distinguished.
 
@@
 The point here is that we would need the appropriate space to go through all of these details, which would probably double the size of the paper as we would have to introduce a lot of previous work that we cannot expect the reader to be familiar with.
-Our purpose therefore is limited to only show that the Borel sets contain a big class of experimentally interesting propositions that are not contained in the QL logic lattice.
-
+Our purpose therefore is limited only to showing that the Borel sets contain a big class of experimentally interesting propositions that are not contained in the QL lattice.
 We added the following footnote to both clarify and give some more information about the issue:
 
-\change{Note that we are not claiming that all Borel sets can be constructed in this way or that all Borel sets correspond to physically interesting propositions.
+\textcolor{blue}{Note that we are not claiming that all Borel sets can be constructed in this way or that all Borel sets correspond to physically interesting propositions.
 First of all, it is not true: $\{ \psi \}$ and $\{ e^{\imath \theta} \psi\}$ are different Borel sets, yet they are not physically distinguishable since a difference in absolute phase is not physically relevant.
 This can be fixed by considering the Borel sets of the projective space.
-Second, to really examine the physicality of all Borel sets we need a ``general theory of experimental logic'', of the type provided by [cite kevin kelly] or [cite carcassi aidala], which would go well beyond the present discussion.
+Second, to really examine the physicality of all Borel sets we need a ``general theory of experimental logic'', of the type provided by \cite{Kelly:1996} or \cite{Carcassi:2021}, which would go well beyond the present discussion.
 With appropriate caveats (e.g.
-the Hilbert space must be separable), one can say that the Borel sets corresponds to propositions that can be associated to an experimental test (what [carcassi aidala] call theoretical statements) which may or may not terminate in any or all cases.
+the Hilbert space must be separable), one can say that the Borel sets correspond to propositions that can be associated to an experimental test (what \cite{Carcassi:2021} call ``theoretical statements'') which may or may not terminate in any or all cases.
 We leave this discussion for another work.}
 
-For (b), TODO
+This footnote responds also to (b).
 
-For (c), each interpretation of quantum mechanics will have a competing view on the meaning of the expectation values and on the meaning of the propositions.
-For example, for some (e.g.
-Qubism, TODO) all expectations will indeed provide only information about the state of knowledge of the agent because, in their view, that is all that quantum states provide.
-For other (e.g.
-Bohmian, TODO) they will represent expectation over hidden variables.
+Finally, for (c) each interpretation of quantum mechanics will have a competing view on the meaning of the expectation values and on the meaning of the propositions.
+For example, for some (e.g.\ Qubism, Relational QM) all expectations will indeed provide only information about the state of knowledge of the agent because, in their view, that is all that quantum states provide.
+For others (e.g.
+Bohmian mechanics, Nelson mechanics) they will represent expectations over hidden variables.
 
-Regardless of the interpretation, the framework of quantum mechanics provides a well-defined expectation value operation for either a  pure state $\psi$ and a density matrix $\rho$.
-Therefore the statement ``quantity A has expectation value r'' will correspond to the mathematical expression $\langle \psi | A | \psi \rangle = r$ (or $\text{Tr}(\rho) = r$).
-It is up to the particular interpretation to make it clear what those expectation mean.
-The logical structure (i.e.
-the Borel sets) will be the same regardless.
+Regardless of the interpretation of QM one may endorse, the framework of quantum mechanics provides a well-defined expectation value operation for either a pure state $\psi$ or a density matrix $\rho$.
+Therefore the statement ``quantity A has expectation value r'' will correspond to the mathematical expression $\langle \psi | A | \psi \rangle = r$ (or $\text{Tr}(\rho A) = r$).
+It is up to the particular interpretation to make it clear what those expectations mean.
+The logical structure (i.e.\ the Borel sets) will be the same regardless.
+\end{enumerate}
 
-TODO: add a footnote somewhere?
+\noindent With these answers we hope to have properly responded to the comments of reviewer 2, which have been very useful to improve the quality of the present essay.
 
---------------------
-
-In closing, we were very pleasantly surprised by the quality of both
-reports and we thank both Referees for the time they spent in
-evaluating our paper.
-
-Sincerely,
-
-The Authors
-
+\clearpage
+\bibliographystyle{apalike}
+\bibliography{bibliography}
 \end{document}

### 2022-04-20 Articles/2021-HarriganSpekkens/Submissions/PRL/response.tex
HarriganSpekkens: draft response

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amssymb}
\usepackage{color}
\usepackage{graphicx}
\usepackage{epsfig,amssymb,amsmath,amsthm}
%\usepackage[active]{srcltx}
%\usepackage[hypertex,linkcolor=red]{hyperref}

\usepackage{color}
\usepackage{cancel}
\usepackage{tikz-cd}

\DeclareMathOperator{\spn}{span}

\newcommand{\pj}[1] {\underbar{$#1$}}
%\newcommand{\pj}[1] {\overline{#1}}

\def\>{\rangle}
\def\<{\langle}
\def\ca{_{\cal A}}
\def\cb{_{\cal B}}
\def\cc{_{\cal C}}
\def\comment#1{}
%\def\comment#1{ [{\bf Comment Lor:} {\sf #1}]}
\def\commentg#1{ [{\bf Comment Gabriele:} {\sf #1}]}
\def\labell#1{\label{#1}}
%\def\labell#1{\label{#1}{\mbox{{\tiny #1}}}}
%\def\section#1{{\par\em #1:--- }}
\def\togli#1{}
\def\sh{\mbox{sh}}
\def\iden{\openone}

\begin{document}

	TODO: thank the editor

	We apologize in advance if we state fact obvious to the reader, but we found that the knowledge on the foundations of probability among our physics colleagues varies greatly, so we'd rather err on the side of caution.

	The overall tone of the response from Barrett is negative. At times he seems to think that we are aiming to contradict the PBR theorem. But in the closing remark, Barret states:
	\begin{quote}
		We know from the PBR theorem that
		any such model must either be $\psi$-ontic or violate preparation independence.
		Possibly there is a result to be proven along the lines suggested above, that
		classical entropies in such a model cannot match quantum entropies.
	\end{quote}
	This is literally what we are doing. In the introduction we say:
	\begin{quote}
		In this paper we are not going to rebut the theorem
		itself.

		...

		Combining these results with PBR, our conclusion is that
		the HS classification itself is fundamentally problematic.
	\end{quote}
	It clearly states that we are using the PBR theorem, not contradicting it. As Barrett says, the PBR theorem rules out a sub-category of ontological models: psi-epistemic models for which preparations are assumed independent. We are not refuting that. What we show is that another sub-category, the psi-ontic models, are to be ruled out as well, as they do not agree with quantum information theory. It is the combination of our result and the PBR result that makes us conclude that the HS categorization is essentially ``empty''.

	The re-evaluation of PBR is not that PBR is wrong: is that many have thought it to mean that psi-ontic states are the answer. We point out that they are not: that the whole HS classification is problematic. The closing remark suggests that our intent is aligned, therefore it is not clear to us where the disagreement is. The result that rule out $\psi$-ontic models is stated in the abstract: ``In this paper we show that the HS framework has a fundamental problem: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics. Namely, \textbf{the map between the epistemic states of the model and quantum density matrices preserves neither the value nor ordering of the information entropy.}''. This point is neither acknowledged, refuted nor commented in Barrett's response.

	We are left very confused.

	Before going into the merit of some of the technical details, let us step back a bit. In an ontological model the epistemic state is represented by the probability $p(\lambda | P)$ of obtaining an ontological state $\lambda$ given a preparation $P$. Classical mixture is treated like classical probability conditioning (i.e. $\sum_i  w_i p(\lambda| P_{\psi_i})$ - eq.~3 in our paper). This is a classical (Kolmogorov) probability density (given that the space $\Lambda$ has to have at least cardinality of the continuum), meaning that the probability space has to satisfy the standard probability axioms (e.g. countable additivity). So the question is really: can a classical (Kolmogorov, countably additive) probability space replicate all the results of a quantum probability space?

	Since Bell, this question has been answered by finding specific cuts on the probability, to show that this or that model does not work. We are using a different strategy. The foundations of both classical probability and classical information theory lies within (classical) measure theory. The results on one side are linked to the results on the other side: if we have a uniform distribution $\rho$ with support $U$, we have $S(\rho) = \log \mu(U)$. Since the function is invertible, violations in information theory are necessarily violation of probability theory and vice-versa. On the other hand, on physics grounds, violations of information theory by themselves are enough to break statistical mechanics and thermodynamics predictions, and therefore would correspond to physical violations.

	On the quantum side, the foundations of quantum probability lies not in classical measures, rather in the inner product of Hilbert spaces. The inner product is also the foundations for quantum information theory and, like in the classical case, the two are linked. Suppose we have two pure states $\rho_\psi$ and $\rho_\phi$, the entropy $S(\frac{1}{2} \rho_\psi + \frac{1}{2} \rho_\phi)$ is given by the Shannon entropy $S(\frac{1+p}{2}, \frac{1-p}{2})$ where $p=\<\psi | \phi \>\<\phi | \psi\>$. Again, this is an invertible function, so violations of probability are violation of information theory.

	Therefore, we can rephrase the question above as: can classical information theory replicate all the results of quantum information theory? We already know that quantum information theory violates bounds given by classical information theory. For example, a quantum composite state can have information entropy lower than the individual systems, something that classical information theory cannot replicate. Classical information theory is understood to be a ``strict subset'' of quantum information theory, and therefore, intuitively, we already expect this not to work. Using this strategy has also a technical advantage in that the space of classical and quantum mixtures are easy to compare: they are both a linear space, where the linearity has the same physical meaning (i.e. statistical mixture). Classical and quantum probability spaces, on the other hand, are rather different as we lay out in the discussion section of the paper. The difference is so profound that Von Neumann and others thought (erroneously in our view) that the rules of logic had to be rewritten. So, going the information theory route bypasses all those problems.

	Now, given this premise, as we say in the paper:
	\begin{quote}
		... we will concentrate on a single aspect of the map: how information entropy transforms under $\iota$. While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is
		sufficient to show that the epistemic structures of $E_HS$
		and $E_QM$ are different
	\end{quote}
	That is, since we have to show that classical information theory cannot replicate quantum information theory, we only need to find one particular feature that is irremediably broken. We settled on showing that mixing break the ordering of the information entropy since ordering is the most fundamental property of any quantitative system: the notion of high and low information gets lost. This is a deadly blow which, again, should not be surprising as we already know that quantum information theory allows us to do things that classical information theory doesn't (it wouldn't be such an interesting field of research otherwise). This is shown in Fig.~2, when one sees, for example, the third case and the second case switching places.

	Now, whether this different way to frame our result helps or not, we leave it to the reviewers. If it does, we can, of course, significantly alter the paper. It will help, though, to address Barrett's comments.

	Barrett focuses on two details. The first is that the map $\iota : E_HS \to E_QS$ from epistemic states in the HS model to quantum mixed states is not invertible. 
	\begin{quote}
The paper points out that in this model, the different mixtures
correspond to different distributions over $\lambda$, even though they yield the
same quantum mixed state I/2.

Although the paper does not do this, one can in fact ask: must *any* ontological
model for a qubit (not just the $\psi$-complete model considered) have the feature
that different mixtures corresponding to the same quantum mixed state correspond
to different distributions over $\lambda$? In fact, the answer is known to be yes,
and this property of quantum systems is known as `preparation contextuality’.
This was shown in R.W.Spekkens, Phys. Rev. A 71, 052108 (2005), the first paper
to introduce preparation contextuality, as well as in many other papers since.
	\end{quote}

	The fact that the map is not invertible is not a problem per-se. As we say in the paper, the question is:
	\begin{quote}
The key question, then, is the following: is the
map $\iota$ sufficiently well behaved that we can understand
the epistemic content of $E_{QM}$ based on the epistemic
content of $E_{HS}$?
	\end{quote}
We contend that it is not, given that the ordering is broken as Fig.~2 shows, and this claim does not seem to us to be either acknowledged or challenged.

As for `preparation contextuality', the question is: can we still use classical conditioning of the form $\sum_i  w_i p(\lambda| P_i)$? To our understanding, that is the case. Therefore we are still in a classical (Kolmogorov) space. The rules of how distributions and information entropy combine are the same. The argument stands.

The second comment focuses on the infinite vs finite value for entropy. The only objection to the failure is:
\begin{quote}
Again, though, the conclusions here are only reached via consideration of the psi-complete model.
\end{quote}
Let us address that objection. Suppose we have a $\psi$-incomplete model. An epistemic state will be a weighted sum $\sum_i  w_i p(\lambda| P_{\psi_i})$ of non overlapping functions $p(\lambda| P_{\psi_i})$. This means that the information entropy is given by $S(w_i) + \sum_i S(p(\lambda| P_{\psi_i}))$. Now, since the information entropy for each pure state must be zero to agree with quantum information, $S(p(\lambda| P_{\psi_i}))=0$ for all $i$, the entropy of the mixture reduces to the Shannon entropy, which reduces to the $\psi$-complete case. This cover all $\psi$-ontic cases.

For $\psi$-epistemic, we already know from PBR that those with preparation independence violate quantum probability, therefore (by the arguments before) will also violate quantum information. Now consider a quantum state for which you have independence at the quantum probability level. The quantum probability does factorize meaning that the quantum information over the composite is the sum of the information over the individual systems. In classical probability, the information of the joint distribution is the sum of the information over the marginals if and only if the joint distribution is the product of the marginals. Therefore quantum information forces you to work in the case of independence, which was ruled out on the grounds that it does not respect quantum probability.

Barrett then asks
\begin{quote}
A better question would be whether there even exists a
(necessarily preparation-contextual) ontological model for a qubit, in which the
Shannon (or differential) entropy of the $\lambda$-distribution is always equal to
the von Neumann entropy of a corresponding mixed state, at least up to an
additive constant. 
\end{quote}
Well, this is literally what we are proving: the order is broken, what is greater on one side may or may not be greater on the other side. A real number is a type of order (one that is complete, dense, with no end-points and contains a countable dense subset), if ordering is broken, real numbers are broken.

Barrett comments:
\begin{quote}
This way, the quantum entropy actually would reflect the
classical ignorance about $\lambda$.
\end{quote}
Whether we can interpret the difference in entropy as a classical ignorance is debated in the paper:
\begin{quote}
Ideally, one would like to say that $E_HS$ represents
all possible epistemic states that include the knowledge
about preparation, while $E_QM$ represents only those distinguishable
through measurement.
\end{quote}
And we argue that the breaking of the ordering is exactly why can't we do that. No comment is given on that argument.

No other further comments are given to the rest of the paper, where the full conclusion is taken, the root of the problem is found in the different nature of quantum probability spaces, a technical definition of quantum contextuality is given and it is shown how it cannot be expressed with a single classical (Kolmogorov) measure.

Barrett also mentions:
\begin{quote}
(Technical remark: in a sense it is infinite, but it would be better to say
undefined. A proper treatment of entropies for distributions over continuous
variables requires differential entropies.)
\end{quote}
This gives a chance to go more into the technical details. In a measure theoretic sense, the measure is literally infinite, not undefined. The image of a measure is the extended real line, which means $\mathbb{R} \cup \{-\infty, +\infty\}$. The question of whether one uses the Shannon entropy or the differential entropy is exactly whether one uses a counting measure or the standard measure on $\mathbb{R}^n$. Very often, mathematical details are not important for physics, but in this case this detail is crucial, and in fact it's the very mathematical detail which makes quantum mechanics/probability/information different.

In classical measure theory, we need to choose a measure which, physically, defines how we ``count'' states in physics. We can either choose the counting measure, so that the measure of a set, it's size, is $\mu(U) = \#(U)$, where we basically just count the number of points. So, for each point $x$ we have $\mu(\{x\}) = 1$. Each point is associated with finite information, each point identifies one case. This measure is suitable for finite (or countable) cases (e.g. the number of faces on a die, the number of boxes a ball can be placed in). Naturally, any set with infinite points will have infinite measure (again, literally infinite) and therefore infinite entropy.

In classical measure theory, on a real line, we can choose the Lebesgue measure. The measure is the size of the interval: $\mu([a, b]) = b-a$. In this case, finite intervals are associated with finite measure, but point will necessarily have zero measure. Finite intervals are associated with finite information, while points are associated with $-\infty$ information.

The point is: in classical measure theory you have to make a choice. If one chooses the Lebesque measure, one has to use probability densities and the differential entropy. If one chooses the counting measure, one has to use proper probabilities and Shannon entropy. By using the theory of distributions, we can blur the lines by using things like the delta function to properly talk about finite probability on a point over a continuum, but the differential entropy of a delta Dirac is $-\infty$, consistent with the fact that a point has measure zero.

We ``hint'' at this technical imprecision in our footnote:
\begin{quote}
	We will follow the convention of not distinguishing between probability and probability density.
\end{quote}
We do believe this is bad, because it is hiding the actual problem. But such is the practice in our field, and so we adapt.

Now, quantum measure theory breaks all of this because both points and open regions have an entropy that would correspond to finite measure. That is, each pure state has zero entropy, which would correspond to measure one. A uniform distribution over the whole Bloch sphere has entropy of 1 which would correspond to measure $2^1 = 2$. You see: a single point has only double the measure of a finite area, which has infinitely many points.

If we were to recover a measure corresponding to the sets over which we defined uniform distributions, the measure would not be additive. Quite literally: $1+1\neq 2$. Quantum mechanics breaks measure theory and it easy to see that it must. Consider the following requirements:
\begin{enumerate}
	\item each state (point) has measure one
	\item each finite region has finite measure
	\item the measure is additive
\end{enumerate}
They are clearly incompatible since a finite region will have uncountably many points. The counting measure gets rid of requirement 2. The Lebesgue measure gets rid of requirement 1. Quantum mechanics gets rid of requirement 3. Classical probability is recovered in quantum mechanics over subsets (i.e. orthogonal basis) where the measure is additive.

This also address the closing question:
\begin{quote}
But the problem then becomes,
how do we describe nature in a way that does not fit the schema of ontological
models?	
\end{quote}
We need to go back and redo the foundations of measure theory such as we can construct a generalized non-additive version, that allows both classical and quantum mechanics as a special case. This is, naturally, well outside the scope of the present article, which just wants to show that ontological models are not the way to do it.

We hope that the discussion addresses at least some of the technical questions. It is not clear to us exactly how to address them in the paper, as we feel the main criticism didn't touch the actual point of the paper (the breakdown of ordering between classical and quantum information entropy), and some of the questions raised are already addressed in the paper.

Sincerely,

The Authors

\end{document}

### 2022-04-20 Articles/2021-HarriganSpekkens/Submissions/PRL/response.tex
HarriganSpekkens: Andrea's modifications

@@
 
-	TODO: thank the editor
+\title{\textbf{Replies to Professor Barrett Comments}}
 
-	We apologize in advance if we state fact obvious to the reader, but we found that the knowledge on the foundations of probability among our physics colleagues varies greatly, so we'd rather err on the side of caution.
+\author{The Authors}
 
-	The overall tone of the response from Barrett is negative. At times he seems to think that we are aiming to contradict the PBR theorem. But in the closing remark, Barret states:
+\maketitle
+
+\textcolor{blue}{We would like to thank the Editor for the opportunity to respond to Professor Barrett engaging comments about our manuscript, these remarks not only allow us to explain in more detail our main result, but also to say a few words on its theoretical foundations.
+We hope to have properly addressed his doubts and objections with the following text; similarly, we hope that these answers will be sufficient in order to send the paper to the next step of the review process.}
+
+\textcolor{blue}{Before stating our replies to Barrett's comments, we apologize in advance if we will mention facts that may sound obvious to some reader; however, since the knowledge on the foundations of probability among our physics colleagues varies greatly, we would rather err on the side of caution.}
+\vspace{2mm}
+
+\textcolor{blue}{The overall tone of Barrett response is negative, at times he seems to think that we are aiming to contradict the PBR theorem. However, in the closing remark he writes that:}
 \begin{quote}
-		We know from the PBR theorem that
-		any such model must either be $\psi$-ontic or violate preparation independence.
-		Possibly there is a result to be proven along the lines suggested above, that
-		classical entropies in such a model cannot match quantum entropies.
+We know from the PBR theorem that	 any such model must either be $\psi$-ontic or violate preparation independence.
+Possibly there is a result to be proven along the lines suggested above, that classical entropies in such a model cannot match quantum entropies.
 \end{quote}
-	This is literally what we are doing. In the introduction we say:
+
+\noindent \textcolor{blue}{Interestingly, this is literally what we aim to show in our essay; in fact, in the introduction we say:}
 \begin{quote}		
-		In this paper we are not going to rebut the theorem
-		itself.
+In this paper we are not going to rebut the theorem itself.
 
@@
 
-		Combining these results with PBR, our conclusion is that
-		the HS classification itself is fundamentally problematic.
+Combining these results with PBR, our conclusion is that the HS classification itself is fundamentally problematic.
 \end{quote}
-	It clearly states that we are using the PBR theorem, not contradicting it.
-As Barrett says, the PBR theorem rules out a sub-category of ontological models: psi-epistemic models for which preparations are assumed independent.
-We are not refuting that.
-What we show is that another sub-category, the psi-ontic models, are to be ruled out as well, as they do not agree with quantum information theory.
-It is the combination of our result and the PBR result that makes us conclude that the HS categorization is essentially ``empty''.
 
-	The re-evaluation of PBR is not that PBR is wrong: is that many have thought it to mean that psi-ontic states are the answer.
-We point out that they are not: that the whole HS classification is problematic.
-The closing remark suggests that our intent is aligned, therefore it is not clear to us where the disagreement is.
-The result that rule out $\psi$-ontic models is stated in the abstract: ``In this paper we show that the HS framework has a fundamental problem: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics.
-Namely, \textbf{the map between the epistemic states of the model and quantum density matrices preserves neither the value nor ordering of the information entropy.}''.
-This point is neither acknowledged, refuted nor commented in Barrett's response.
+	\textcolor{blue}{This statement clearly affirms that in our argument we are using the PBR theorem, not contradicting it.
+As Barrett correctly says, the PBR result rules out a sub-category of ontological models: $\psi$-epistemic models for which preparations are assumed independent.
+We are \emph{not} refuting that claim.
+However, many scholars think that the PBR theorem proves that $\psi$-ontic states are the correct interpretation of quantum states, we point out that such a view is not correct by showing that the whole HS classification is problematic.
+Indeed, what we show is that another sub-category, the $\psi$-ontic models, are to be ruled out as well, since they do not agree with quantum information theory.
+It is the combination of our result and the PBR theorem that makes us conclude that the HS categorization is essentially ``empty''.}
 
-	We are left very confused.
+		\textcolor{blue}{The above mentioned closing remark suggests that our intent is aligned, therefore, it is not clear to us where the disagreement is.} The result that rule out $\psi$-ontic models is stated in the abstract:
+\begin{quote}
+In this paper we show that the HS framework has a fundamental problem: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics.
+Namely, \textbf{the map between the epistemic states of the model and quantum density matrices preserves neither the value nor ordering of the information entropy.}
+\end{quote}
+\noindent This point is neither acknowledged, refuted nor commented in Barrett's response, this fact left us very confused.
 
-	Before going into the merit of some of the technical details, let us step back a bit.
+	Before going into the merit of some of the technical details, let us take a step back.
 In an ontological model the epistemic state is represented by the probability $p(\lambda | P)$ of obtaining an ontological state $\lambda$ given a preparation $P$.
 Classical mixture is treated like classical probability conditioning (i.e.
-$\sum_i  w_i p(\lambda| P_{\psi_i})$ - eq.~3 in our paper).
-This is a classical (Kolmogorov) probability density (given that the space $\Lambda$ has to have at least cardinality of the continuum), meaning that the probability space has to satisfy the standard probability axioms (e.g.
-countable additivity).
-So the question is really: can a classical (Kolmogorov, countably additive) probability space replicate all the results of a quantum probability space?
+$\sum_i  w_i p(\lambda| P_{\psi_i})$, cf.\ eq.~3 in our paper).
+This is a classical (Kolmogorov) probability density (given that the space $\Lambda$ must have cardinality of the continuum at least), meaning that the probability space has to satisfy the standard probability axioms (e.g.\ countable additivity).
+\textcolor{blue}{Hence, the crucial question is: can a classical (Kolmogorov, countably additive) probability space replicate all the results of a quantum probability space?}
 
 	Since Bell, this question has been answered by finding specific cuts on the probability, to show that this or that model does not work.
-We are using a different strategy.
-The foundations of both classical probability and classical information theory lies within (classical) measure theory.
+We are employing a different strategy.
+The foundations of both classical probability and classical information theory lie within (classical) measure theory.
 The results on one side are linked to the results on the other side: if we have a uniform distribution $\rho$ with support $U$, we have $S(\rho) = \log \mu(U)$.
 Since the function is invertible, violations in information theory are necessarily violation of probability theory and vice-versa.
-On the other hand, on physics grounds, violations of information theory by themselves are enough to break statistical mechanics and thermodynamics predictions, and therefore would correspond to physical violations.
+On the other hand, on physical grounds, violations of information theory by themselves are enough to break statistical mechanics and thermodynamics predictions, and therefore would correspond to physical violations.
 
-	On the quantum side, the foundations of quantum probability lies not in classical measures, rather in the inner product of Hilbert spaces.
-The inner product is also the foundations for quantum information theory and, like in the classical case, the two are linked.
-Suppose we have two pure states $\rho_\psi$ and $\rho_\phi$, the entropy $S(\frac{1}{2} \rho_\psi + \frac{1}{2} \rho_\phi)$ is given by the Shannon entropy $S(\frac{1+p}{2}, \frac{1-p}{2})$ where $p=\<\psi | \phi \>\<\phi | \psi\>$.
+	\textcolor{blue}{Taking into account quantum mechanics it is well-known that the foundations of quantum probability have not to be found in classical measures, rather in the inner product of Hilbert spaces, which also constitutes the foundations for quantum information theory and, like in the classical case, the two are linked.} Suppose we have two pure states $\rho_\psi$ and $\rho_\phi$, the entropy $S(\frac{1}{2} \rho_\psi + \frac{1}{2} \rho_\phi)$ is given by the Shannon entropy $S(\frac{1+p}{2}, \frac{1-p}{2})$ where $p=\<\psi | \phi \>\<\phi | \psi\>$.
 Again, this is an invertible function, so violations of probability are violation of information theory.
 
-	Therefore, we can rephrase the question above as: can classical information theory replicate all the results of quantum information theory?
-We already know that quantum information theory violates bounds given by classical information theory.
+	Therefore, we can rephrase the question above as follows: can classical information theory replicate all the results of quantum information theory?
+We already know that the latter violates bounds given by the former.
 For example, a quantum composite state can have information entropy lower than the individual systems, something that classical information theory cannot replicate.
@@
 Classical and quantum probability spaces, on the other hand, are rather different as we lay out in the discussion section of the paper.
-The difference is so profound that Von Neumann and others thought (erroneously in our view) that the rules of logic had to be rewritten.
-So, going the information theory route bypasses all those problems.
+\textcolor{blue}{The difference is so profound that von Neumann and others thought (erroneously in our view) that the rules of classical propositional logic had to be modified.
+Thus, going the information theory route bypasses all those problems.}
+	\vspace{2mm}
 
@@
 While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is
-		sufficient to show that the epistemic structures of $E_HS$
-		and $E_QM$ are different
+		sufficient to show that the epistemic structures of $E_{HS}$
+		and $E_{QM}$ are different
 	\end{quote}
@@
 We settled on showing that mixing break the ordering of the information entropy since ordering is the most fundamental property of any quantitative system: the notion of high and low information gets lost.
-This is a deadly blow which, again, should not be surprising as we already know that quantum information theory allows us to do things that classical information theory doesn't (it wouldn't be such an interesting field of research otherwise).
-This is shown in Fig.~2, when one sees, for example, the third case and the second case switching places.
+This is a deadly blow which, again, should not be surprising as we already know that quantum information theory allows us to do things that classical information theory does not (it would not be such an interesting field of research otherwise).
+This fact is shown in Fig.~2, when one sees, for example, the third case and the second case switching places.
 
@@
 
-	Barrett focuses on two details. The first is that the map $\iota : E_HS \to E_QS$ from epistemic states in the HS model to quantum mixed states is not invertible. 
+	Barrett focuses on two details. The first is that the map $\iota : E_{HS} \to E_{QS}$ from epistemic states in the HS model to quantum mixed states is not invertible. 
 	\begin{quote}
-The paper points out that in this model, the different mixtures
-correspond to different distributions over $\lambda$, even though they yield the
-same quantum mixed state I/2.
+The paper points out that in this model, the different mixtures correspond to different distributions over $\lambda$, even though they yield the same quantum mixed state I/2.
 
-Although the paper does not do this, one can in fact ask: must *any* ontological
-model for a qubit (not just the $\psi$-complete model considered) have the feature
-that different mixtures corresponding to the same quantum mixed state correspond
-to different distributions over $\lambda$? In fact, the answer is known to be yes,
-and this property of quantum systems is known as `preparation contextuality’.
-This was shown in R.W.Spekkens, Phys. Rev. A 71, 052108 (2005), the first paper
-to introduce preparation contextuality, as well as in many other papers since.
+Although the paper does not do this, one can in fact ask: must *any* ontological model for a qubit (not just the $\psi$-complete model considered) have the feature that different mixtures corresponding to the same quantum mixed state correspond to different distributions over $\lambda$?
+In fact, the answer is known to be yes, and this property of quantum systems is known as `preparation contextuality’.
+This was shown in R.W.\ Spekkens, Phys.
+Rev.
+A 71, 052108 (2005), the first paper to introduce preparation contextuality, as well as in many other papers since.
 	\end{quote}
 
-	The fact that the map is not invertible is not a problem per-se. As we say in the paper, the question is:
+	The fact that the map is not invertible is not a problem per se. As we say in the paper, the question is:
 	\begin{quote}
@@
 To our understanding, that is the case.
-Therefore we are still in a classical (Kolmogorov) space.
+Therefore, we are still in a classical (Kolmogorov) space.
 The rules of how distributions and information entropy combine are the same.
@@
 
-The second comment focuses on the infinite vs finite value for entropy. The only objection to the failure is:
+The second comment focuses on the infinite versus finite value for entropy. The only objection to the failure is:
 \begin{quote}
-Again, though, the conclusions here are only reached via consideration of the psi-complete model.
+Again, though, the conclusions here are only reached via consideration of the $\psi$-complete model.
 \end{quote}
@@
 Suppose we have a $\psi$-incomplete model.
-An epistemic state will be a weighted sum $\sum_i  w_i p(\lambda| P_{\psi_i})$ of non overlapping functions $p(\lambda| P_{\psi_i})$.
+An epistemic (\textcolor{blue}{ontic?}) state will be a weighted sum $\sum_i  w_i p(\lambda| P_{\psi_i})$ of non overlapping functions $p(\lambda| P_{\psi_i})$.
 This means that the information entropy is given by $S(w_i) + \sum_i S(p(\lambda| P_{\psi_i}))$.
@@
 Now consider a quantum state for which you have independence at the quantum probability level.
-The quantum probability does factorize meaning that the quantum information over the composite is the sum of the information over the individual systems.
+The quantum probability does factorize, meaning that the quantum information over the composite is the sum of the information over the individual systems.
 In classical probability, the information of the joint distribution is the sum of the information over the marginals if and only if the joint distribution is the product of the marginals.
-Therefore quantum information forces you to work in the case of independence, which was ruled out on the grounds that it does not respect quantum probability.
+Therefore, quantum information forces one to work in the case of independence % mi puoi chiarire meglio questa frase, non capisco bene il punto.
+which was ruled out on the grounds that it does not respect quantum probability.
 
-Barrett then asks
+Barrett then asks:
 \begin{quote}
@@
 \end{quote}
-Well, this is literally what we are proving: the order is broken, what is greater on one side may or may not be greater on the other side.
+\textcolor{blue}{Remarkably, this is literally what we are proving}: the order is broken, what is greater on one side may or may not be greater on the other side.
 A real number is a type of order (one that is complete, dense, with no end-points and contains a countable dense subset), if ordering is broken, real numbers are broken.
@@
 \begin{quote}
-Ideally, one would like to say that $E_HS$ represents
-all possible epistemic states that include the knowledge
-about preparation, while $E_QM$ represents only those distinguishable
-through measurement.
+Ideally, one would like to say that $E_{HS}$ represents all possible epistemic states that include the knowledge about preparation, while $E_{QM}$ represents only those distinguishable through measurement.
 \end{quote}
-And we argue that the breaking of the ordering is exactly why can't we do that. No comment is given on that argument.
+And we argue that the breaking of the ordering is exactly why we cannot do that. Nonetheless, no comment is given on that argument.
 
-No other further comments are given to the rest of the paper, where the full conclusion is taken, the root of the problem is found in the different nature of quantum probability spaces, a technical definition of quantum contextuality is given and it is shown how it cannot be expressed with a single classical (Kolmogorov) measure.
+\textcolor{blue}{Along the same lines---and quite surprisingly---no further comments are given to the rest of the paper}, where the full conclusion is taken, the root of the problem is found in the different nature of quantum probability spaces, a technical definition of quantum contextuality is given, and it is shown how it cannot be expressed with a single classical (Kolmogorov) measure.
 
-Barrett also mentions:
+\textcolor{blue}{Barrett also mentions a technical remark concerning Shannon entropy:}
 \begin{quote}
-(Technical remark: in a sense it is infinite, but it would be better to say
-undefined. A proper treatment of entropies for distributions over continuous
-variables requires differential entropies.)
+in a sense it is infinite, but it would be better to say undefined. A proper treatment of entropies for distributions over continuous variables requires differential entropies.
 \end{quote}
-This gives a chance to go more into the technical details.
+
+This gives a chance to go more into the technical details of our work.
 In a measure theoretic sense, the measure is literally infinite, not undefined.
@@
 The question of whether one uses the Shannon entropy or the differential entropy is exactly whether one uses a counting measure or the standard measure on $\mathbb{R}^n$.
-Very often, mathematical details are not important for physics, but in this case this detail is crucial, and in fact it's the very mathematical detail which makes quantum mechanics/probability/information different.
+\textcolor{blue}{Very often, mathematical details are not important for physics, but in the present case this detail is crucial, and in fact it is the very mathematical detail which makes quantum mechanics, quantum probability and quantum information theory different with respect to their classical counterparts.}
 
-In classical measure theory, we need to choose a measure which, physically, defines how we ``count'' states in physics.
+In classical measure theory, we need to choose a measure which defines how we ``count'' states in physics.
 We can either choose the counting measure, so that the measure of a set, it's size, is $\mu(U) = \#(U)$, where we basically just count the number of points.
@@
 A uniform distribution over the whole Bloch sphere has entropy of 1 which would correspond to measure $2^1 = 2$.
-You see: a single point has only double the measure of a finite area, which has infinitely many points.
+Hence, a single point has only double the measure of a finite area, which has infinitely many points.
 
-If we were to recover a measure corresponding to the sets over which we defined uniform distributions, the measure would not be additive.
-Quite literally: $1+1\neq 2$.
-Quantum mechanics breaks measure theory and it easy to see that it must.
-Consider the following requirements:
+If we were to recover a measure corresponding to the sets over which we defined uniform distributions, the measure would not be additive. %Quite literally: $1+1\neq 2$. 
+Quantum mechanics breaks measure theory, and it easy to see that it must. Consider the following requirements:
 \begin{enumerate}
@@
 \end{quote}
-We need to go back and redo the foundations of measure theory such as we can construct a generalized non-additive version, that allows both classical and quantum mechanics as a special case.
+We need to go back and redo the foundations of measure theory such as we can construct a generalized non-additive version of it, that allows both classical and quantum mechanics as a special case.
 This is, naturally, well outside the scope of the present article, which just wants to show that ontological models are not the way to do it.
 
-We hope that the discussion addresses at least some of the technical questions.
-It is not clear to us exactly how to address them in the paper, as we feel the main criticism didn't touch the actual point of the paper (the breakdown of ordering between classical and quantum information entropy), and some of the questions raised are already addressed in the paper.
+We hope that our discussion addresses at least some of the most important technical questions raised by Barrett'c comments.
+It is not clear to us exactly how to address them in the paper, as we feel the main criticism contained in his report did not touch the actual point of the paper (the breakdown of ordering between classical and quantum information entropy), and some of his questions are already addressed in the paper.
+\vspace{5mm}
 
 Sincerely,
+\vspace{5mm}
 

### 2022-04-20 Articles/2021-HarriganSpekkens/Submissions/PRL/response.tex
HarriganSpekkens: reviewed response with Christine

@@
 
-\title{\textbf{Replies to Professor Barrett Comments}}
+\title{\textbf{Reply to Professor Barrett's Comments}}
 
@@
 
-\textcolor{blue}{We would like to thank the Editor for the opportunity to respond to Professor Barrett engaging comments about our manuscript, these remarks not only allow us to explain in more detail our main result, but also to say a few words on its theoretical foundations.
-We hope to have properly addressed his doubts and objections with the following text; similarly, we hope that these answers will be sufficient in order to send the paper to the next step of the review process.}
+We would like to thank the Editor for the opportunity to respond to Professor Barrett's engaging comments about our manuscript, these remarks not only allow us to explain in more detail our main result, but also to say a few words on its theoretical foundations.
+We hope to have properly addressed his doubts with the following text; similarly, we hope that these answers will be sufficient in order to send the paper to the next step of the review process.
 
-\textcolor{blue}{Before stating our replies to Barrett's comments, we apologize in advance if we will mention facts that may sound obvious to some reader; however, since the knowledge on the foundations of probability among our physics colleagues varies greatly, we would rather err on the side of caution.}
+We apologize in advance if we mention facts that may sound obvious to some readers; however, since knowledge of the the foundations of probability among our physics colleagues varies greatly, we would rather err on the side of caution.
 \vspace{2mm}
 
-\textcolor{blue}{The overall tone of Barrett response is negative, at times he seems to think that we are aiming to contradict the PBR theorem. However, in the closing remark he writes that:}
+There seems to be some sort of misunderstanding or disconnect between our intent and Barrett's perspective on the manuscript.
+In particular, he seems to think that we are aiming to contradict the PBR theorem.
+But, in the closing remark, he adds that:
 \begin{quote}
@@
 
-\noindent \textcolor{blue}{Interestingly, this is literally what we aim to show in our essay; in fact, in the introduction we say:}
+Ironically, this is literally what we aim to show in our paper; in fact, in the introduction we say:
 \begin{quote}		
@@
 
-	\textcolor{blue}{This statement clearly affirms that in our argument we are using the PBR theorem, not contradicting it.
+This statement clearly affirms that in our argument we are using the PBR theorem, not contradicting it.
 As Barrett correctly says, the PBR result rules out a sub-category of ontological models: $\psi$-epistemic models for which preparations are assumed independent.
 We are \emph{not} refuting that claim.
-However, many scholars think that the PBR theorem proves that $\psi$-ontic states are the correct interpretation of quantum states, we point out that such a view is not correct by showing that the whole HS classification is problematic.
+However, many scholars think that the PBR theorem proves that $\psi$-ontic states are the correct interpretation of quantum states.
+We point out that such a view is not correct by showing that the whole HS classification is problematic.
 Indeed, what we show is that another sub-category, the $\psi$-ontic models, are to be ruled out as well, since they do not agree with quantum information theory.
-It is the combination of our result and the PBR theorem that makes us conclude that the HS categorization is essentially ``empty''.}
+It is the combination of our result and the PBR theorem that makes us conclude that the HS categorization is essentially ``empty''.
 
-		\textcolor{blue}{The above mentioned closing remark suggests that our intent is aligned, therefore, it is not clear to us where the disagreement is.} The result that rule out $\psi$-ontic models is stated in the abstract:
+The above mentioned closing remark suggests that our intent is aligned with his; therefore, it is not clear to us where the disagreement is.
+The result that rules out $\psi$-ontic models is stated in the abstract:
 \begin{quote}
@@
 \end{quote}
-\noindent This point is neither acknowledged, refuted nor commented in Barrett's response, this fact left us very confused.
+\noindent This point is not acknowledged, refuted or commented on in Barrett's response; this fact left us very confused.
 
@@
 This is a classical (Kolmogorov) probability density (given that the space $\Lambda$ must have cardinality of the continuum at least), meaning that the probability space has to satisfy the standard probability axioms (e.g.\ countable additivity).
-\textcolor{blue}{Hence, the crucial question is: can a classical (Kolmogorov, countably additive) probability space replicate all the results of a quantum probability space?}
+Hence, the crucial question is: can a classical (Kolmogorov, countably additive) probability space replicate all the results of a quantum probability space?
 
-	Since Bell, this question has been answered by finding specific cuts on the probability, to show that this or that model does not work.
+	Since Bell, this question has been answered by finding specific inequalities/constraints on the probability, to show that this or that model does not work.
 We are employing a different strategy.
@@
 The results on one side are linked to the results on the other side: if we have a uniform distribution $\rho$ with support $U$, we have $S(\rho) = \log \mu(U)$.
-Since the function is invertible, violations in information theory are necessarily violation of probability theory and vice-versa.
-On the other hand, on physical grounds, violations of information theory by themselves are enough to break statistical mechanics and thermodynamics predictions, and therefore would correspond to physical violations.
+Since the function is invertible, constraints on information theory are necessarily constraints on probability theory and vice-versa.
+Therefore violations of the former coincide with violations of the latter.
+Moreover, on physical grounds, violations of information theory by themselves are enough to break statistical mechanics and thermodynamics predictions, and therefore would correspond to physical violations.
 
-	\textcolor{blue}{Taking into account quantum mechanics it is well-known that the foundations of quantum probability have not to be found in classical measures, rather in the inner product of Hilbert spaces, which also constitutes the foundations for quantum information theory and, like in the classical case, the two are linked.} Suppose we have two pure states $\rho_\psi$ and $\rho_\phi$, the entropy $S(\frac{1}{2} \rho_\psi + \frac{1}{2} \rho_\phi)$ is given by the Shannon entropy $S(\frac{1+p}{2}, \frac{1-p}{2})$ where $p=\<\psi | \phi \>\<\phi | \psi\>$.
+	On the other hand, the foundations of quantum probability are not based on classical measures, rather on the inner product of Hilbert spaces, which also constitutes the foundations for quantum information theory.
+As in the classical case, the two are linked.
+Suppose we have two pure states $\rho_\psi$ and $\rho_\phi$ expressed as density matrices.
+The entropy $S(\frac{1}{2} \rho_\psi + \frac{1}{2} \rho_\phi)$ is given by the Shannon entropy $S(\frac{1+p}{2}, \frac{1-p}{2})$ where $p=\<\psi | \phi \>\<\phi | \psi\>$.
 Again, this is an invertible function, so violations of probability are violation of information theory.
@@
 For example, a quantum composite state can have information entropy lower than the individual systems, something that classical information theory cannot replicate.
-Classical information theory is understood to be a ``strict subset'' of quantum information theory, and therefore, intuitively, we already expect this not to work.
-Using this strategy has also a technical advantage in that the space of classical and quantum mixtures are easy to compare: they are both a linear space, where the linearity has the same physical meaning (i.e.
+Classical information theory is understood to be a ``strict subset'' of quantum information theory, and therefore, intuitively, we already expect that classical information theory cannot replicate all the results of quantum information theory.
+Using this strategy has also a technical advantage in that the spaces of classical and quantum mixtures are easy to compare: they are both linear spaces, where the linearity has the same physical meaning (i.e.
 statistical mixture).
 Classical and quantum probability spaces, on the other hand, are rather different as we lay out in the discussion section of the paper.
-\textcolor{blue}{The difference is so profound that von Neumann and others thought (erroneously in our view) that the rules of classical propositional logic had to be modified.
-Thus, going the information theory route bypasses all those problems.}
+The difference is so profound that von Neumann and others thought (erroneously in our view) that the rules of classical propositional logic had to be modified.
+Thus, going the information theory route bypasses all those problems.
 	\vspace{2mm}
@@
 	That is, since we have to show that classical information theory cannot replicate quantum information theory, we only need to find one particular feature that is irremediably broken.
-We settled on showing that mixing break the ordering of the information entropy since ordering is the most fundamental property of any quantitative system: the notion of high and low information gets lost.
+We settled on showing that mixing breaks the ordering of the information entropy since ordering is the most fundamental property of any quantitative system: the notion of greater and lesser information gets lost.
 This is a deadly blow which, again, should not be surprising as we already know that quantum information theory allows us to do things that classical information theory does not (it would not be such an interesting field of research otherwise).
-This fact is shown in Fig.~2, when one sees, for example, the third case and the second case switching places.
+The breakdown of ordering is shown in Fig.~2, where one sees, for example, the third case and the second case switching places.
 
 	Now, whether this different way to frame our result helps or not, we leave it to the reviewers.
-If it does, we can, of course, significantly alter the paper.
+If it does, we can, of course, significantly alter the text of the paper.
 It will help, though, to address Barrett's comments.
@@
 
-The second comment focuses on the infinite versus finite value for entropy. The only objection to the failure is:
+The second comment focuses on the infinite versus finite value for entropy. The only objection to the claimed failure is:
 \begin{quote}
@@
 Suppose we have a $\psi$-incomplete model.
-An epistemic (\textcolor{blue}{ontic?}) state will be a weighted sum $\sum_i  w_i p(\lambda| P_{\psi_i})$ of non overlapping functions $p(\lambda| P_{\psi_i})$.
-This means that the information entropy is given by $S(w_i) + \sum_i S(p(\lambda| P_{\psi_i}))$.
-Now, since the information entropy for each pure state must be zero to agree with quantum information, $S(p(\lambda| P_{\psi_i}))=0$ for all $i$, the entropy of the mixture reduces to the Shannon entropy, which reduces to the $\psi$-complete case.
-This cover all $\psi$-ontic cases.
+In this case, an epistemic state will be a weighted sum $\sum_i  w_i p(\lambda| P_{\psi_i})$ of \textbf{nonoverlapping} functions $p(\lambda| P_{\psi_i})$.
+This means that the information entropy of the mixture reduces to $S(w_i) + \sum_i S(p(\lambda| P_{\psi_i}))$.
+Now, since the information entropy for each pure state must be zero to agree with quantum information, $S(p(\lambda| P_{\psi_i}))=0$ for all $i$, the entropy of the mixture further reduces to the Shannon entropy $S(w_i)$, which coincides with that of the $\psi$-complete case.
+This means that all we say for the $\psi$-complete models equally applies to all $\psi$-incomplete models, and therefore to all $\psi$-ontic cases.
 
-For $\psi$-epistemic, we already know from PBR that those with preparation independence violate quantum probability, therefore (by the arguments before) will also violate quantum information.
-Now consider a quantum state for which you have independence at the quantum probability level.
-The quantum probability does factorize, meaning that the quantum information over the composite is the sum of the information over the individual systems.
+While our result is directed toward $\psi$-ontic models, we can construct another argument for the remaining $\psi$-epistemic ones.
+We already know from PBR that the models that assume preparation independence violate quantum probability, therefore (by the arguments before) will also violate quantum information.
+Now consider a quantum state for which you have independence at the quantum probability level, that is the density matrix factorizes.
+In this case, the quantum information over the composite is the sum of the information over the individual systems.
 In classical probability, the information of the joint distribution is the sum of the information over the marginals if and only if the joint distribution is the product of the marginals.
-Therefore, quantum information forces one to work in the case of independence % mi puoi chiarire meglio questa frase, non capisco bene il punto.
-which was ruled out on the grounds that it does not respect quantum probability.
+Therefore, on one side, if we want quantum information to agree with classical information, the probabilities must factorize, but on the other side PBR tells us that if the probabilities factorize then we have a violation of quantum probabilities.
+This is a contradiction, so we can't replicate both quantum probabilities and quantum information at the same time.
+To our understanding, this means that all $\psi$-epistemic models must be ruled out.
+
+We concede that we haven't explored these points in the manuscript. We would be happy to add them if the reviewers think they are valuable.
 
@@
 \end{quote}
-\textcolor{blue}{Remarkably, this is literally what we are proving}: the order is broken, what is greater on one side may or may not be greater on the other side.
-A real number is a type of order (one that is complete, dense, with no end-points and contains a countable dense subset), if ordering is broken, real numbers are broken.
+We are puzzled by this comment as we are literally proving that no such model exists.
+The order is broken: a greater von Neumann entropy may or may not correspond to a greater Shannon entropy.
+A real number is a type of order (one that is complete, dense, with no end-points and contains a countable dense subset).
+If ordering is broken, real numbers are broken.
 
@@
 
-\textcolor{blue}{Along the same lines---and quite surprisingly---no further comments are given to the rest of the paper}, where the full conclusion is taken, the root of the problem is found in the different nature of quantum probability spaces, a technical definition of quantum contextuality is given, and it is shown how it cannot be expressed with a single classical (Kolmogorov) measure.
+Along the same lines---and quite surprisingly to us---no further comments are given on the rest of the paper, where the full conclusion is taken, the root of the problem is found in the different nature of quantum probability spaces, a technical definition of quantum contextuality is given, and it is shown how it cannot be expressed with a single classical (Kolmogorov) measure.
 
-\textcolor{blue}{Barrett also mentions a technical remark concerning Shannon entropy:}
+Barrett also mentions a technical remark concerning Shannon entropy:
 \begin{quote}
@@
 
-This gives a chance to go more into the technical details of our work.
+This gives us a chance to go more into the technical details of our work.
 In a measure theoretic sense, the measure is literally infinite, not undefined.
@@
 The question of whether one uses the Shannon entropy or the differential entropy is exactly whether one uses a counting measure or the standard measure on $\mathbb{R}^n$.
-\textcolor{blue}{Very often, mathematical details are not important for physics, but in the present case this detail is crucial, and in fact it is the very mathematical detail which makes quantum mechanics, quantum probability and quantum information theory different with respect to their classical counterparts.}
+Very often, mathematical details are not important for physics, but in the present case this detail is crucial, and in fact it is the very mathematical detail which makes quantum mechanics, quantum probability and quantum information theory different with respect to their classical counterparts.
 
 In classical measure theory, we need to choose a measure which defines how we ``count'' states in physics.
-We can either choose the counting measure, so that the measure of a set, it's size, is $\mu(U) = \#(U)$, where we basically just count the number of points.
+We can choose the counting measure, so that the measure of a set, its size, is $\mu(U) = \#(U)$, where we basically just count the number of points.
 So, for each point $x$ we have $\mu(\{x\}) = 1$.
@@
 
-In classical measure theory, on a real line, we can choose the Lebesgue measure.
+On a real line we can alternatively choose the Lebesgue measure.
 The measure is the size of the interval: $\mu([a, b]) = b-a$.
-In this case, finite intervals are associated with finite measure, but point will necessarily have zero measure.
+In this case, finite intervals are associated with finite measure, but points will necessarily have zero measure.
 Finite intervals are associated with finite information, while points are associated with $-\infty$ information.
@@
 The point is: in classical measure theory you have to make a choice.
-If one chooses the Lebesque measure, one has to use probability densities and the differential entropy.
-If one chooses the counting measure, one has to use proper probabilities and Shannon entropy.
+If one chooses the Lebesgue measure, one has to use probability densities and the differential entropy.
+If one chooses the counting measure, one has to use ``proper'' probabilities and Shannon entropy.
 By using the theory of distributions, we can blur the lines by using things like the delta function to properly talk about finite probability on a point over a continuum, but the differential entropy of a delta Dirac is $-\infty$, consistent with the fact that a point has measure zero.
@@
 A uniform distribution over the whole Bloch sphere has entropy of 1 which would correspond to measure $2^1 = 2$.
-Hence, a single point has only double the measure of a finite area, which has infinitely many points.
+Hence, a single point has only half the measure of a finite area, which has infinitely many points.
 
 If we were to recover a measure corresponding to the sets over which we defined uniform distributions, the measure would not be additive. %Quite literally: $1+1\neq 2$. 
-Quantum mechanics breaks measure theory, and it easy to see that it must. Consider the following requirements:
+Quantum mechanics breaks measure theory, and it is easy to see that it must. Consider the following requirements:
 \begin{enumerate}
@@
 Quantum mechanics gets rid of requirement 3.
-Classical probability is recovered in quantum mechanics over subsets (i.e.
-orthogonal basis) where the measure is additive.
+Classical probability can only be recovered in quantum mechanics over special subsets where the measure happens to be additive (i.e.
+orthogonal basis), though even in those cases, it is only finitely additive in general.
 
-This also address the closing question:
+This also addresses the closing question:
 \begin{quote}
@@
 \end{quote}
-We need to go back and redo the foundations of measure theory such as we can construct a generalized non-additive version of it, that allows both classical and quantum mechanics as a special case.
+We need to go back and redo the foundations of measure theory such that we can construct a generalized non-additive version of it, which allows both classical and quantum mechanics as special cases.
 This is, naturally, well outside the scope of the present article, which just wants to show that ontological models are not the way to do it.
 
-We hope that our discussion addresses at least some of the most important technical questions raised by Barrett'c comments.
+We hope that our discussion addresses at least some of the most important technical questions raised by Barrett's comments.
 It is not clear to us exactly how to address them in the paper, as we feel the main criticism contained in his report did not touch the actual point of the paper (the breakdown of ordering between classical and quantum information entropy), and some of his questions are already addressed in the paper.
+We will look to the referee reports for guidance.
 \vspace{5mm}
@@
 Sincerely,
-\vspace{5mm}
 

### 2022-06-08 Articles/2021-HarriganSpekkens/HS-ShortNote.tex
HarriganSpekkens: polished math part of the note

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{enumitem}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{tikz}

\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\newcommand\partitle[1]{\textsc{#1}.}

\begin{document}

\title{A no go theorem for $\psi$-ontic models}
\author{Gabriele Carcassi}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\author{Andrea Oldofredi}
\affiliation{Centre of Philosophy, University of Lisbon, Portugal}
\author{Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{}

\begin{abstract}
In this note we give a simple argument as to why $\psi$-ontic models cannot be expected to reproduce quantum mechanics. We will see that, using information theoretic considerations, the lack of overlap of epistemic state requires all states to be orthogonal, which openly contradicts quantum theory. This argument is a much simplified version of the one we presented in a previous paper.
\end{abstract}

\maketitle

\section{Introduction}

\section{The argument}

What we want to show is that a set of non-overlapping probability distributions cannot reproduce the entropy given by quantum mechanics. Let us review a couple of properties of the information entropy, both in the classical setting (i.e. the Shannon/Gibbs entropy) and in the quantum setting (i.e. the von Neumann entropy).\footnote{We leave the calculations in the appendix.}

\textbf{Entropy of mixed non-overlapping distributions}. Let $\rho_1$ and $\rho_2$ be two classical distributions over a space $\Lambda$ with measure $\lambda$. The entropy will be given by:
\begin{equation}\label{shannon_entropy}
	H(\rho_1) = - \int_\Lambda \rho_1(\lambda) \log \rho_1(\lambda) d\lambda.
\end{equation}
Suppose the two distributions are disjoint, and let $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ be a uniform mixture of the two distributions. The entropy of $\rho$ is given by
\begin{equation}\label{entropy_nonoverlap}
	H(\rho) = 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2).
\end{equation}
Note how the non-overlapping assumption constrains the entropy of the mixed state.

\textbf{Entropy of quantum mixed states}. Now suppose $\psi$ and $\phi$ are two pure quantum states and let $p = | \langle \psi | \phi \rangle |^2$ the probability of transition from one to the other. Consider the mixed state $\rho = \frac{1}{2} | \psi \rangle \langle \psi | + \frac{1}{2} | \phi \rangle \langle \phi |$. Its entropy is given by
\begin{equation}\label{entropy_mixed}
	H(\rho) = H\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right).
\end{equation}

\textbf{Non-overlapping distributions can only represent orthogonal states}. Now, suppose we have a $\psi$-ontic model. The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping distribution over a space $\Lambda$, therefore eq. \ref{entropy_nonoverlap} applies. Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states is zero, we must have
\begin{equation}\label{entropy_pure}
	H(p(\lambda|P_\psi)) = H(p(\lambda|P_\phi)) = 0,
\end{equation}
and therefore
\begin{equation}\label{required_entropy}
	H\left(p(\lambda|\frac{1}{2}P_\psi + \frac{1}{2}P_\phi)\right) = 1.
\end{equation}
If we compare the above with eq. \ref{entropy_mixed}, it follows that $p$ must be zero. That is, we must have that
\begin{equation}\label{orthogonal}
	 \langle \psi | \phi \rangle = 0
\end{equation}
no matter what $\psi$ and $\phi$.

The non-overlapping assumption built into the $\psi$-ontic model necessarily implies that all pure states are orthogonal. Since this is not true in quantum mechanics, any $\psi$-ontic models will fail to reproduce the results of quantum information, quantum statistical mechanics and, therefore, quantum theory in general.

In retrospect this should not be surprising for two reasons. First of all, the case where all states are orthogonal is exactly the case for which all observables commutes, the classical case. Therefore we are simply finding that classical probability pushes us toward the classical case. Secondly, we already know that quantum information theory has features that are not reproducible in classical information theory, and that is exactly what makes it such an exciting new field.

\section{Conclusion}

We have presented a much simpler version of the argument presented in [cite] to reach the same conclusion: $\psi$-ontic models are not compatible with quantum information theory and therefore quantum theory itself. The finding is consistent with our claim that the ontological model has a hidden assumption of non-contextuality (through the use of standard Kolmogorov probability), which here can be seen to force all pure states to be orthogonal, to belong to the same context, as they do in classical mechanics.

One may think the problem could be circumvented by substituting the Shannon entropy formula. That is, the problem is not with the non-overlapping assumption but rather with how the entropy is calculated. However, the link between probability theory, information theory and measure theory is so tight that it makes this approach unfeasible: failure of one means failure of all. We will look deeper into this issue in a later work.

\section*{Appendix: calculation}

\textbf{Entropy of mixed non-overlapping distributions}. We want to show that, given two non-overlapping probability distributions $\rho_1$ and $\rho_2$, the entropy of $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ is given by
\begin{equation}\label{entropy_nonoverlap_appendix}
	H(\rho) = 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2).
\end{equation}

Let $U_1, U_2 \subset \Lambda$ be the respective support of the distributions. Since the distribution are non-overlapping, we have $U_1 \cap U_2 = \emptyset$. We have
\begin{align*}
	H(\rho) &= - \int_\Lambda \rho \log \rho d\lambda \\
	&= -\int_{U_1} \rho \log \rho d\lambda -\int_{U_2} \rho \log \rho d\lambda \\
	&= -\int_{U_1} \frac{1}{2} \rho_1 \log \frac{1}{2} \rho_1 d\lambda -\int_{U_2} \frac{1}{2} \rho_2 \log \frac{1}{2} \rho_2 d\lambda \\
	&= - \frac{1}{2} \int_{U_1} \rho_1 \log \frac{1}{2} d\lambda - \frac{1}{2} \int_{U_1} \rho_1 \log \rho_1 d\lambda \\
	&- \frac{1}{2} \int_{U_2} \rho_2 \log \frac{1}{2} d\lambda - \frac{1}{2} \int_{U_2} \rho_2 \log \rho_2 d\lambda \\
	&= - \frac{1}{2} \log \frac{1}{2} - \frac{1}{2} \log \frac{1}{2} + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2) \\
	&= 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2) \\
\end{align*}

\textbf{Entropy of quantum mixed states}. We want to show that, given two states $\psi$ and $\phi$, the entropy of the mixed state $\rho = \frac{1}{2}|\psi\rangle\langle\psi| + \frac{1}{2}|\phi\rangle\langle\phi|$ is:
\begin{equation}\label{entropy}
	H(\rho) = H\left(\frac{1+|\langle\psi|\phi\rangle|}{2}, \frac{1-|\langle\psi|\phi\rangle|}{2}\right).
\end{equation}

\begin{center}
	\begin{tikzpicture}[scale = 1]
		\draw (0,0) circle (2);
		\node at (-2.3,0) {$-$};
		\node at (2.3,0) {$+$};
		\node at (2,1.4) {$\psi$};
		\node at (2,-1.4) {$\phi$};
		\draw (-2,0) -- (2,0);
		\begin{scope}
			\clip(0,0) circle (2);
			\draw (0,0) -- (2,1.4);
			\draw (0,0) -- (2,-1.4);
			\draw (1.634,1.4) -- (1.634,-1.4);
		\end{scope}
		\fill (1.634,0) circle (0.05);
		\node at (1.8,.2) {$\rho$};
	\end{tikzpicture}
\end{center}

Note that $\psi$ and $\phi$ will identify a two dimensional subspace which can be though, without loss of generality, as a qubit and therefore can be represented by a Bloch sphere. The picture represents the intersection of the Block sphere with the plane identified by $\psi$ and $\phi$. As $\rho$ is an equal mixture of the two states, it will be represent by the midpoint between the two. Taking the line that goes through $\rho$ and the center of the sphere, we can see that $\rho$ can also be seen as the mixture of the states $+$ and $-$ which, since they represent equal and opposite directions, they are form a basis. To diagonalize $\rho$, then, means to express it in terms of $+$ and $-$.

If $\theta_{\psi\phi}$ is the angle between $\psi$ and $\phi$, we have
\begin{equation}
	|\langle \psi | \phi \rangle |^2 = \cos^2 \frac{\theta_{\psi\phi}}{2}.
\end{equation}
The angle is divided by two because the angle on the Bloch sphere (i.e. in physical space) is double the angle in the Hilber space. For example, for $z^+$ and $z^-$ the angle would be $\pi/2$ and the inner product is zero.

Now we express $\psi$ and $\phi$ in terms of $+$ and $-$, remembering that they form a basis. Given that $\rho$ is on the midpoint, the figure is vertically symmetric. The angle between $\psi$ and $+$, then, is half of $\theta_{\psi\phi}$. We have:
\begin{align*}
	\left|\psi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>+\sin\frac{\theta_{\psi\phi}}{4}\left|-\right> \\
	\left|\phi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>-\sin\frac{\theta_{\psi\phi}}{4}\left|-\right>
\end{align*}

The density matrices corresponding to the pure states are:
\begin{align*}
	\left|\psi\right>\left<\psi\right|&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right|\\
	&+\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right| \\
	\left|\phi\right>\left<\phi\right|&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right|\\
	&-\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right|
\end{align*}

We can now calculate the mixture:
\begin{align*}
	\frac{1}{2}(|\psi\rangle\langle\psi| &+ |\phi\rangle\langle\phi|) \\
	&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right| +\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right| \\
	&=\frac{1+\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|+\right>\left<+\right| +\frac{1-\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|-\right>\left<-\right| \\
	&=\frac{1+|\langle\psi|\phi\rangle|}{2}\left|+\right>\left<+\right| +\frac{1-|\langle\psi|\phi\rangle|}{2}\left|-\right>\left<-\right| \\
\end{align*}

As $\rho$ is in a diagonal form, the entropy is given by \ref{entropy}

\bibliography{bibliography}

\end{document}

### 2022-06-09 Articles/2021-HarriganSpekkens/HS-ShortNote.tex
HarriganSpekkens: reviewed math with Christine

@@
 
-\title{A no go theorem for $\psi$-ontic models}
+\title{A no-go theorem for $\psi$-ontic models}
 \author{Gabriele Carcassi}
@@
 In this note we give a simple argument as to why $\psi$-ontic models cannot be expected to reproduce quantum mechanics.
-We will see that, using information theoretic considerations, the lack of overlap of epistemic state requires all states to be orthogonal, which openly contradicts quantum theory.
+We will see that, using information theoretic considerations, the lack of overlap of epistemic states requires all states to be orthogonal, which directly contradicts quantum theory.
 This argument is a much simplified version of the one we presented in a previous paper.
@@
 
-\textbf{Entropy of mixed non-overlapping distributions}. Let $\rho_1$ and $\rho_2$ be two classical distributions over a space $\Lambda$ with measure $\lambda$. The entropy will be given by:
+\textbf{Entropy of mixed non-overlapping distributions}.
+Let $\rho_1$ and $\rho_2$ be two classical distributions over a space $\Lambda$ with measure $\lambda$.
+The entropy will be given by\footnote{The logarithm is assumed to be in base 2.}
 \begin{equation}\label{shannon_entropy}
@@
 \end{equation}
-Note how the non-overlapping assumption constrains the entropy of the mixed state.
+Note how the non-overlapping assumption fixes the entropy of the mixed state.
 
 \textbf{Entropy of quantum mixed states}.
-Now suppose $\psi$ and $\phi$ are two pure quantum states and let $p = | \langle \psi | \phi \rangle |^2$ the probability of transition from one to the other.
+Now suppose $\psi$ and $\phi$ are two pure quantum states and let $p = | \langle \psi | \phi \rangle |^2$ be the probability of transition from one to the other.
 Consider the mixed state $\rho = \frac{1}{2} | \psi \rangle \langle \psi | + \frac{1}{2} | \phi \rangle \langle \phi |$.
@@
 Now, suppose we have a $\psi$-ontic model.
-The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping distribution over a space $\Lambda$, therefore eq.
+The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping distributions over a space $\Lambda$, therefore eq.
 \ref{entropy_nonoverlap} applies.
@@
 \begin{equation}\label{required_entropy}
-	H\left(p(\lambda|\frac{1}{2}P_\psi + \frac{1}{2}P_\phi)\right) = 1.
+	\begin{aligned}
+	H\left(p(\lambda|\frac{1}{2}P_\psi + \frac{1}{2}P_\phi)\right) &= \\
+H\left(\frac{1}{2}p(\lambda|P_\psi) + \frac{1}{2}p(\lambda|P_\phi)\right) 
+&= 1.
+	\end{aligned}
 \end{equation}
@@
 \end{equation}
-no matter what $\psi$ and $\phi$.
+no matter what $\psi$ and $\phi$ are.
 
@@
 In retrospect this should not be surprising for two reasons.
-First of all, the case where all states are orthogonal is exactly the case for which all observables commutes, the classical case.
-Therefore we are simply finding that classical probability pushes us toward the classical case.
+First of all, the case where all states are orthogonal is exactly the case for which all observables commute, the classical case.
+Therefore we are simply finding that the implicit use of classical probability in $\psi$-ontic models pushes us to the classical case.
 Secondly, we already know that quantum information theory has features that are not reproducible in classical information theory, and that is exactly what makes it such an exciting new field.
@@
 
-Let $U_1, U_2 \subset \Lambda$ be the respective support of the distributions. Since the distribution are non-overlapping, we have $U_1 \cap U_2 = \emptyset$. We have
+Let $U_1, U_2 \subset \Lambda$ be the respective supports of the distributions. Since the distributions are non-overlapping, we have $U_1 \cap U_2 = \emptyset$. We have
 \begin{align*}
@@
 \textbf{Entropy of quantum mixed states}.
-We want to show that, given two states $\psi$ and $\phi$, the entropy of the mixed state $\rho = \frac{1}{2}|\psi\rangle\langle\psi| + \frac{1}{2}|\phi\rangle\langle\phi|$ is:
+We want to show that, given two states $\psi$ and $\phi$, the entropy of the mixed state $\rho = \frac{1}{2}|\psi\rangle\langle\psi| + \frac{1}{2}|\phi\rangle\langle\phi|$ is
 \begin{equation}\label{entropy}
@@
 
-Note that $\psi$ and $\phi$ will identify a two dimensional subspace which can be though, without loss of generality, as a qubit and therefore can be represented by a Bloch sphere.
-The picture represents the intersection of the Block sphere with the plane identified by $\psi$ and $\phi$.
+Note that $\psi$ and $\phi$ will identify a two-dimensional subspace which can be thought, without loss of generality, as a qubit and therefore can be represented by a Bloch sphere.
+The picture represents the intersection of the Bloch sphere with the plane identified by $\psi$ and $\phi$.
 As $\rho$ is an equal mixture of the two states, it will be represent by the midpoint between the two.
-Taking the line that goes through $\rho$ and the center of the sphere, we can see that $\rho$ can also be seen as the mixture of the states $+$ and $-$ which, since they represent equal and opposite directions, they are form a basis.
+Taking the line that goes through $\rho$ and the center of the sphere, we can see that $\rho$ can also be seen as the mixture of the states $+$ and $-$ which, since they represent equal and opposite directions, form a basis.
 To diagonalize $\rho$, then, means to express it in terms of $+$ and $-$.
@@
 The angle is divided by two because the angle on the Bloch sphere (i.e.
-in physical space) is double the angle in the Hilber space.
-For example, for $z^+$ and $z^-$ the angle would be $\pi/2$ and the inner product is zero.
+in physical space) is double the angle in the Hilbert space.
+For example, for $z^+$ and $z^-$ the angle on the Bloch sphere would be $\pi$ and the inner product is zero (i.e.
+opposite directions in physical space correspond to orthogonal states).
 
 Now we express $\psi$ and $\phi$ in terms of $+$ and $-$, remembering that they form a basis.
-Given that $\rho$ is on the midpoint, the figure is vertically symmetric.
+Given that $\rho$ is at the midpoint, the figure is vertically symmetric.
 The angle between $\psi$ and $+$, then, is half of $\theta_{\psi\phi}$.
-We have:
+The inner product between $\psi$ and $+$ is
+\begin{equation}
+	\begin{aligned}
+	|\langle \psi | + \rangle |^2 &= \cos^2 \frac{\theta_{\psi +}}{2} \\
+&= \cos^2 \frac{\theta_{\psi\phi}}{4}.
+	\end{aligned}
+\end{equation}
+Keeping in mind that we are composing vectors in the Hilbert space (and not in the geometry of the physical space) we have
 \begin{align*}
 	\left|\psi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>+\sin\frac{\theta_{\psi\phi}}{4}\left|-\right> \\
-	\left|\phi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>-\sin\frac{\theta_{\psi\phi}}{4}\left|-\right>
+	\left|\phi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>-\sin\frac{\theta_{\psi\phi}}{4}\left|-\right>.
 \end{align*}
 
-The density matrices corresponding to the pure states are:
+The density matrices corresponding to the pure states are
 \begin{align*}
@@
 	&-\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
-	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right|
+	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right|.
 \end{align*}
 
-We can now calculate the mixture:
+We can now calculate the mixture
 \begin{align*}
@@
 	&=\frac{1+\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|+\right>\left<+\right| +\frac{1-\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|-\right>\left<-\right| \\
-	&=\frac{1+|\langle\psi|\phi\rangle|}{2}\left|+\right>\left<+\right| +\frac{1-|\langle\psi|\phi\rangle|}{2}\left|-\right>\left<-\right| \\
+	&=\frac{1+|\langle\psi|\phi\rangle|}{2}\left|+\right>\left<+\right| +\frac{1-|\langle\psi|\phi\rangle|}{2}\left|-\right>\left<-\right|. \\
 \end{align*}
 
-As $\rho$ is in a diagonal form, the entropy is given by \ref{entropy}
+As $\rho$ is in a diagonal form, the entropy is given by eq.~\ref{entropy}.
 

### 2022-06-10 Articles/2021-HarriganSpekkens/HS-ShortNote.tex
HarriganSpekkens: review after Andrea

@@
 
-\title{A no-go theorem for $\psi$-ontic models}
+\title{A No-Go Theorem for $\psi$-Ontic Models}
 \author{Gabriele Carcassi}
@@
 
-\date{}
+\date{9 June 2022}
 
 \begin{abstract}
-In this note we give a simple argument as to why $\psi$-ontic models cannot be expected to reproduce quantum mechanics.
-We will see that, using information theoretic considerations, the lack of overlap of epistemic states requires all states to be orthogonal, which directly contradicts quantum theory.
-This argument is a much simplified version of the one we presented in a previous paper.
+In this note we give a simple no-go theorem that explains why $\psi$-ontic models cannot be expected to reproduce quantum mechanics.
+We will see that, using information theoretic considerations, the lack of overlap of epistemic states requires all states to be orthogonal, which openly contradicts quantum theory.
+This argument is a simplified version of the one we presented in a \href{https://arxiv.org/pdf/2201.11842.pdf}{previous paper}.
 \end{abstract}
@@
 \section{Introduction}
+In 2010 Harrigan and Spekkens (HS) proposed a formal classification to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property instantiated by a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}.
+Although the original aim of this classification was to clarify Einstein's view about quantum mechanics (QM), the HS framework has been widely employed not only to categorize different interpretations of QM, but also to argue what formulations of the theory are admissible (\cite{PBR:2012, Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Lewis:2012, Ladyman:2021} for critical discussions).
+
+Given the influence of the HS framework in quantum foundations, and in particular the prominent role played by $\psi$-ontic models in several arguments supporting the reality of the quantum state---many of them built upon \cite{PBR:2012}---we re-examine the definitions and features of such models in order to understand whether they constitute a sound basis from which to draw conclusions on the nature of $\psi$.
+Our analysis suggests that they are not.
+
+Referring to this, in this note we provide a no-go theorem for $\psi$-ontic models which is a simplified version of the main result contained in our previous paper \cite{Carcassi:2022}.
+More precisely, here we show that, using information theoretic considerations, $\psi$-ontic models cannot reproduce QM since the lack of overlap of epistemic states requires all states to be orthogonal, which openly contradicts quantum theory.
+Therefore, we conclude, they should not be employed in order to classify quantum models.\footnote{We assume that the reader is familiar with HS definitions.}
 
@@
 
-What we want to show is that a set of non-overlapping probability distributions cannot reproduce the entropy given by quantum mechanics.
-Let us review a couple of properties of the information entropy, both in the classical setting (i.e.
-the Shannon/Gibbs entropy) and in the quantum setting (i.e.
-the von Neumann entropy).\footnote{We leave the calculations in the appendix.}
+We aim to show that a set of non-overlapping probability distributions cannot reproduce the entropy given by quantum mechanics.
+In order to show this claim, let us review a property of the information entropy, both in the classical setting (i.e.\ the Shannon/Gibbs entropy) and in the quantum setting (i.e.\ the von Neumann entropy).\footnote{We show the explicit calculations in the appendix.}
 
@@
 Let $\rho_1$ and $\rho_2$ be two classical distributions over a space $\Lambda$ with measure $\lambda$.
-The entropy will be given by\footnote{The logarithm is assumed to be in base 2.}
+The entropy for each distribution is given by the usual formula\footnote{The logarithm is assumed to be in base 2.}, for example
 \begin{equation}\label{shannon_entropy}
@@
 
-\textbf{Non-overlapping distributions can only represent orthogonal states}.
-Now, suppose we have a $\psi$-ontic model.
+\emph{\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}}
+
+\emph{Proof}: Suppose we have a $\psi$-ontic model defined according to \cite{Harrigan:2010}.
 The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping distributions over a space $\Lambda$, therefore eq.
@@
 \end{equation}
-If we compare the above with eq. \ref{entropy_mixed}, it follows that $p$ must be zero. That is, we must have that
+If we compare the above with eq.~\ref{entropy_mixed}, it follows that $p$ must be zero. That is, we must have that
 \begin{equation}\label{orthogonal}
@@
 
-The non-overlapping assumption built into the $\psi$-ontic model necessarily implies that all pure states are orthogonal.
-Since this is not true in quantum mechanics, any $\psi$-ontic models will fail to reproduce the results of quantum information, quantum statistical mechanics and, therefore, quantum theory in general.
+Hence, the non-overlapping assumption built into the $\psi$-ontic model necessarily implies that all pure states are orthogonal.
+Since this is not true in quantum mechanics, any $\psi$-ontic model will fail to reproduce the results of quantum information, quantum statistical mechanics and, therefore, quantum theory in general.
+$\square$
 
 In retrospect this should not be surprising for two reasons.
-First of all, the case where all states are orthogonal is exactly the case for which all observables commute, the classical case.
-Therefore we are simply finding that the implicit use of classical probability in $\psi$-ontic models pushes us to the classical case.
+First of all, the case where all states are orthogonal corresponds exactly to the case in which all observables commute, i.e.~the classical case.
+Therefore, we are simply finding that the implicit use of classical probability in $\psi$-ontic models pushes us to classical interpretation of quantum models.
 Secondly, we already know that quantum information theory has features that are not reproducible in classical information theory, and that is exactly what makes it such an exciting new field.
@@
 
-We have presented a much simpler version of the argument presented in [cite] to reach the same conclusion: $\psi$-ontic models are not compatible with quantum information theory and therefore quantum theory itself.
-The finding is consistent with our claim that the ontological model has a hidden assumption of non-contextuality (through the use of standard Kolmogorov probability), which here can be seen to force all pure states to be orthogonal, to belong to the same context, as they do in classical mechanics.
+We have presented a simpler and more compact version of the main result presented in \cite{Carcassi:2022}.
+Both arguments reach the same conclusion: $\psi$-ontic models are not compatible with quantum information theory, and therefore with quantum mechanics itself.
 
-One may think the problem could be circumvented by substituting the Shannon entropy formula.
+One may think the issue could be circumvented by substituting the Shannon entropy formula.
 That is, the problem is not with the non-overlapping assumption but rather with how the entropy is calculated.
-However, the link between probability theory, information theory and measure theory is so tight that it makes this approach unfeasible: failure of one means failure of all.
+However, the link between probability theory, information theory and measure theory is so tight that it makes this approach unfeasible.
+If we take a set of states $U$ with a measure $\mu(U)$ that quantifies the states in the set, the probability of each state for a uniform distribution is given by $\rho = 1 / \mu(U)$ and the entropy by $\log \mu(U)$.
+The measure used to count states, the probabilities assigned to states and the entropy assigned to distributions are not independent mathematical structures: failure of one means failure of all.
 We will look deeper into this issue in a later work.
 
+\bibliography{bibliography}
+\clearpage
+
 \section*{Appendix: calculation}
-
+\label{A}
 \textbf{Entropy of mixed non-overlapping distributions}.
 We want to show that, given two non-overlapping probability distributions $\rho_1$ and $\rho_2$, the entropy of $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ is given by
-\begin{equation}\label{entropy_nonoverlap_appendix}
+\begin{equation}
 	H(\rho) = 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2).
@@
 	&= - \frac{1}{2} \log \frac{1}{2} - \frac{1}{2} \log \frac{1}{2} + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2) \\
-	&= 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2) \\
+	&= 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2). \\
 \end{align*}
@@
 
-\bibliography{bibliography}
-
 \end{document}

### 2022-07-03 TeachingMaterial/ActionPrinciple/ActionPrinciple.tex
ActionPrinciple: minor edits

@@
 We give a geometrical interpretation for the action principle in Lagrangian classical particle mechanics that also has a direct physical meaning.
-The idea is that the variation of the action along a path is the flow of the displacement field through the area between the path and the variation.
-Since actual trajectories are parallel to the displacement field, the flow is zero along them, making the action stationary.
+In a nutshell, the variation of the action between a path and its variation effectively ``counts'' the possible evolutions that ``go through'' the area enclosed.
+If a possible evolution is taken, all neighbouring evolutions will be parallel, making them tangent to the area enclosed by the evolution and its variation, making the action stationary.
 We first give the interpretation for a single degree of freedom using simple differential calculus so that it is accessible to the widest possible audience.
@@
 This allowed us to recast Hamiltonian mechanics as the dynamics of deterministic and reversible evolution, which coincidentally gives us the proper setting to characterize both the physics and the geometry underlying the principle of stationary action.
-We will see that the principle arises as a feature of divergence-free fields and closed two-forms, as a general mathematical property of vector calculus and differential geometry.
+We will see that the principle arises as a feature of divergence-free fields and closed two-forms, as a general mathematical property of vector calculus and differential geometry\cite{souriau1997structure}.
 

### 2022-07-03 Articles/2021-ReversePhysics/ReversePhysics.tex
ReversePhysics: Changing the abstract
Reverse Physics: final changes to archive version

@@
 
-\title{Reverse Physics}
+\title{Reverse Physics: From Laws to Physical Assumptions}
 \author{Gabriele Carcassi, Christine A. Aidala}
@@
 \section{Introduction}
-
-%\begin{itemize}
-%	\item Brief descriptions of the aims of foundations of mathematics
-%	\item We want to follow a similar approach (not new interpretations of QM or new models for quantum gravity)
-%	\item Reverse mathematics start from theorems and finds the smallest set of axioms needed to rederive them.
-We want to take physical theories and find the smalles set of physical assumptions that are equivalent to the theories themselves.
-%	\item This paper is meant to showcase the approach and how it can lead to renewed physical insights into theories that may be thought as already fully understood and those that are not.
-The main point of the paper is not the results per se, but the novel of the approach proposed.
-We hope that other people will see the value and pursue similar lines of investigations.
-%	
-%\end{itemize}
 
@@
 
-%Use "investigation trumps rigor"
-%Somewhere, note that "physics" entropy does not coincide with "information" entropy. Physics entropy is information about states as quantified by invariants.
-
 \section{Determinism, reversibility and Hamiltonian mechanics}
-
-%\begin{itemize}
-%	\item Start with the equations of HM
-%	\item Show Liouville's theorem and wonder what it means. Jacobian is 1. Show how Jacobian for 1 d.o.f. is really a Poisson bracket.
-%	\item Connect to statistical mechanics: volume is number of states. HM conserves the number of states. States are mapped one to one on a continuum. (discuss problems of counting on a continuum)
-%	\item Show that we can go backwards on a single degree of freedom
-%	\item Wonder what to do on multiple d.o.f. 
-%	\item Show symplectic form as the geometrical description. Generalize to d.o.f. by taking outer product. Show that areas remain orthogonal
-%	\item What does orthogonality mean? Independence of d.o.f. (volume product of areas)
-%	\item Show that preservation of the geometry (areas and perpendicularity) under time evolution gives us HM
-%	\item HM is det/rev motion of indepent d.o.f.
-%	\item We used count of states, log count of states is also conserved. What about information entropy? It is also conserved. 
-%	\item Show we can proceed the other way. Det/rev means information before and after is the same. Information needs to be conserved -> Jacobian = 1 -> HM.
-%	\item Note how bijection is not enough. There is an additional constraint. Why?
-%	\item Measurements on a continuum have uncertainty.
-Connect to statistics and use the covariance matrix to describe a gaussian peak.
-The determinant of the matrix with the Jacobian.
-If jacobian is 1, determinant is the same, same precision of the description.
-HM <-> to requiring a bijection at the same level of precision.
-%	\item Wonder why we are getting conservation of energy from det/rev, while we typically associate conservation of energy with system isolation (in the thermodynamic sense).
-Argue that if a system is deterministic and reversible, it means that the future/past state depends only on the system itself.
-The system does not depend on anything else -> it is isolated.
-There is a deep connection between isolation and det/rev
-%	
-%\end{itemize}
 
@@
 
-%\begin{itemize}
-%	\item We saw that classical mechanics is more than just points: it says something about uncertainty as well. Quantum mechanics has an uncertainty principle. Is there any connections?
-%	\item Since there was also a connection to entropy, turn to entropy in quantum mechanics.
-Entropy is bounded from below: all pure states have zero entropy.
-In classical mechanics, however, we can prepare distributions with arbitrarily small entropy.
-Can the fact that the entropy is bound from below in QM be the source of the uncertainty?
-%	\item Assume entropy is bounded from below in CM. Show you get uncertainty relationship in a single degree of freedom.
-%	\item Uncertainty relationship <-> bounded entropy = bounded information
-%\end{itemize}
-
 Let us now turn our attention to quantum mechanics, and see if our approach can shed new light on a theory that is not generally considered to be well understood.
@@
 \section{A new understanding of the third law}
-
-%\begin{itemize}
-%	\item We saw that in quantum mechanics entropy is bounded from below. Why?
-%	\item Turn to thermodynamics.
-The third also says that thermodynamic entropy (not information entropy) is bounded from below.
-The zero thermodynamic entropy, however, is formulated in terms of a pure crystal at zero temperature.
-This does not seem to be a fundamental characterization.
-Can we do better?
-%	\item Note how entropy is an extensive quantity.
-Now consider the empty system.
-Since the composition of any system with the empty system must yield the same system, it also must yield the same entropy.
--> Entropy of empty system must be zero.
-Better more fundamental characterization of zero entropy.
-%	\item What would a system with negative entropy correspond to?
-To a description of a system that is more precise than saying the system is not there.
-But that is absurd!
-Therefore, it is a logical necessity that the entropy of a system is bound from below.
-%\end{itemize}
 
@@
 
-%\section{A more unified physics}
-%
-%\begin{itemize}
-%	\item Put physics back at the center of the discussion (we are exploring physical concepts and their relationships, the math is there to carry out the arguments)
-%	\item Appeal to tradition (thinking about small set of physical assumptions was what foundations of classical mechanics, thermodynamics, ... was about)
-%	\item Conceptual unification vs theoretical unification
-%	\item Common standard tools (like set theory, category theory, ...)
-%	\item Connection between fields
-%	\item Reduce the number of fundamental concepts
-%	\item These more powerful tools may help theoretical unification
-%	\item realm of validity of the different theoretical theories
-%	\item realm of validity of the underlying mathematical tools
-%\end{itemize}
-
 \section{Conclusion}
@@
 We hope others will present their own reverse physics results, and that this may lead to a renewed interest in the manner in which physicists in the past tried to formulate the foundations of our field: through physical principles, laws and assumptions.
+
+\section*{Acknowledgments}
+We acknowledge funding from the MCubed program of the University of Michigan.
+This work is part of a larger project, Assumptions of Physics~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
 

### 2022-07-03 Overviews/Differentiability/Differentiability.tex
Differentiability: strawmen for differentiability as polyline limit

@@
 
+\section{Differentiability as reducibility}
+
+%Given a topological embedding of a closed segment and a sequence into the segment with dense image, let for every $n$ $\lambda_n$ be the polyline defined by the first $n$ elements in the sequence.
+We say that the polyline converges to the embedded if for every open neighbourhood of the image of the embedding there is an $N$ such that all polylines beyond $N$ are in the neighbourhood.
+
+%Let X be a manifold and \psi be a chart (a choice of coordinate {x^i}).
+Let \lambda be a curve and {P_i} be a dense subset \lambda.
+Let \lambda_k be the polyline consisting of line segments (straightness is defined by the coordinates {x^i}) identified by the first k points in {P_i}.
+We say that \lambda is smooth if, for any dense subset the sequence of the \lambda_k convergence to \lambda.
+
+\textbf{Conjecture:} a curve is differentiable if it can be seen as a limit of a polyline approximation that is independent on the details of the approximation.
+
+\textbf{Lines and slopes.} Let $U \subseteq \mathbb{R}^n$.
+Let $L$ be the set of all straight lines.
+Given two points $P, Q \in U$, we note $l(P,Q) \in L$ the line that passes through $P$ and $Q$.
+That is, $\{P, Q\} \subset l(P,Q)$.
+We note $l_1 \parallel l_2$ when two lines $l_1, l_2 \in L$ are parallel.
+Parallelism is an equivalence relationship.
+We call the quotient set $S = L_{/\parallel}$ the set of slopes.
+Given a point $P \in U$ and a slope $s \in S$, we note $l(P, s) \in L$ the line that passes through $P$ with the given slope.
+That is, $P \in l(P, s) \in s$.
+It can be shown that $Q \in l(P, s)$ if and only if $P \in l(Q, s)$.
+Note that the topology of $U$ endows a topology on both $L$ and $Q$ which is Hausdorff and second countable, and therefore $L$ and $Q$ possess a notion of limit.
+
+\textbf{Oriented segments.} An oriented segment (or simply a segment) is a triplet $\langle P, s, Q \rangle$ with $P, Q \in U$ and $s \in S$ such that $P \in l(Q,s)$.
+If $P \neq Q$ the segment is said finite; if not, it is said infinitesimal.
+In can be shown that given two distinct points $P, Q$, there is only one oriented segment $\langle P, s, Q \rangle$.
+It can also be shown that $\langle P, s, P \rangle$ is a segment for all $s \in S$.
+We say a sequence of oriented segments $\{\langle P_i, s_i, Q_i \rangle\}_{i=1}^{\infty}$ converges if the limit $\lim\limits_{i \to \infty} \langle P_i, s_i, Q_i \rangle\ = \langle \lim\limits_{i \to \infty} P_i, \lim\limits_{i \to \infty} s_i, \lim\limits_{i \to \infty} Q_i \rangle$ exists.
+
+\textbf{Polylines and refinement.} Two segments $\langle P_1, s_1, Q_1 \rangle$ and $\langle P_2, s_2, Q_2 \rangle$ are consecutive if $Q_1 = P_2$..
+A polyline is a finite sequence of finitely many consecutive segments.
+That is, a polyline is a sequence $\{\langle P_i, s_i, Q_i \rangle\}_{i=1}^{n}$ such that $Q_i = P_{i+1}$.
+A point $P$ belongs to a polyline if there is some $1 \leq i \leq n$ such that $P=P_i$ or $P=Q_i$.
+A polyline $\{\langle P^a_i, s^a_i, Q^a_i \rangle\}_{i=1}^{n}$ is said to refine another polyline $\{\langle P^b_j, s^b_j, Q^b_j \rangle\}_{j=1}^{m}$ if $n > m$ and all the points that belong to the second belongs to the first as well without changing their order.
+That is, for all $1 \leq j \leq m$ we can find $1 \leq i \leq n$ such that $P^a_i = P^b_j$ and also find $1 \leq k \leq n$ such that $Q^a_k = Q^b_j$.
+Additionally, we say an oriented segment $\langle P^a_i, s^a_i, Q^a_i \rangle$ refines a segment $\langle P^b_j, s^b_j, Q^b_j \rangle$ if we can find $k$ and $l$ such that $k \leq i \leq l$, $P^b_j = P^a_k$ and $Q^b_j = Q^a_l$.
+
+\textbf{Polyline approximating sequence (PAS).} A polyline approximating sequence (PAS) is a sequence of polylines such that each one refines the previous.
+We say a point $P \in U$ belongs to the PAS if it is one of the boundary points of a segment in a polyline of the sequence.
+Given a PAS, a segment approximating sequence (SAS) is a sequence of segments such that the $i$-th segment is a segment of the $i$-th polyline, and each segment in the sequence refines the previous segment in the sequence.
+We say a PAS converges if all the SAS converge.
+
+\textbf{Polyline approximation and differentiability.} Let $\lambda \subset U$ be a curve (i.e.
+a topologically embedded interval).
+A polyline approximation for $\lambda$ is a PAS such that the set of points that belong to the PAS is a dense set of $\lambda$.
+It can be shown that if a SAS of a polyline approximation converges, it does so to an infinitesimal segment which we call limit segment.
+Moreover, it can be shown that there will be a limit segment for each point of $\lambda$.
+We say that a curve is differentiable is the limit segment for each point exists and it the same for all polyline approximations.
+
 \bibliography{bibliography}

### 2022-07-08 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
ActionPrinciple: starting to reframe the note as an article

--- 2022-01-30 TeachingMaterial/ActionPrinciple/ActionPrinciple.tex
@@
 \begin{abstract}
-We give a geometrical interpretation for the action principle in Lagrangian classical particle mechanics that also has a direct physical meaning.
-The idea is that the variation of the action along a path is the flow of the displacement field through the area between the path and the variation.
-Since actual trajectories are parallel to the displacement field, the flow is zero along them, making the action stationary.
-We first give the interpretation for a single degree of freedom using simple differential calculus so that it is accessible to the widest possible audience.
-We then generalize to multiple degrees of freedom using symplectic geometry.
-This treatment makes the geometry of both Hamiltonian and Lagrangian mechanics more readily evident, and gives reasons to privilege the Hamiltonian picture over the Lagrangian one.
+We give a geometrical interpretation for the principle of stationary action in classical Lagrangian particle mechanics.
+In a nutshell, the difference of the action along a path and its variation effectively ``counts'' the possible evolutions that ``go through'' the area enclosed.
+If the path corresponds to a possible evolution, all neighbouring evolutions will be parallel, making them tangent to the area enclosed by the path and its variation, thus yielding a stationary action.
+This treatment gives a full physical account of the geometry of both Hamiltonian and Lagrangian mechanics which is founded on three assumptions: determinism and reversible evolution, independence of the degrees of freedom and equivalence between kinematics and dynamics.
+The logical equivalence between the three assumptions and the principle of stationary action leads to a much cleaner conceptual understanding.
 \end{abstract}
@@
 As part of our larger project Assumptions of Physics, we developed an approach, called Reverse Physics, which examines current theories to find a set of starting physical assumptions that are sufficient to rederive them.
-This allowed us to recast Hamiltonian mechanics as the dynamics of deterministic and reversible evolution, which coincidentally gives us the proper setting to characterize both the physics and the geometry underlying the principle of stationary action.
-We will see that the principle arises as a feature of divergence-free fields and closed two-forms, as a general mathematical property of vector calculus and differential geometry.
+This gives a more physically motivated grounding of the classical theory and the proper setting to characterize both the physics and the geometry underlying the principle of stationary action.
+The assumption of determinism and reversible evolution corresponds to a diverge-free flow of the states, which preserves volumes in phase space (i.e.
+the count of states).
+The assumption of independence of the degrees of freedom means the volume reduces to the product of areas in each d.o.f., which is quantified by a closed two-form.
+The principle of stationary action arises as a feature of divergence-free fields and closed two-forms, as a general mathematical property of vector calculus and differential geometry\cite{souriau1997structure}.
+The additional assumption of equivalence between kinematics and dynamics (i.e.
+we can reconstruct the dynamical state simply by looking at the trajectory) is what allows us to express the principle in Lagrangian form.
+The argument can proceed in the reverse direction: assuming the principle of stationary action recovers a dynamical system that exhibit those three physical assumptions.
 
-We will first address the case of a single degree of freedom using standard vector calculus.
+The mathematics needed to run the argument is not new, yet it is typically presented in a way and in a notation that is opaque to physicists.
+To provide a clear insight to the widest possible readership, we will first address the case of a single degree of freedom using standard vector calculus.
 This provides already all the geometrical and physical insights in a language that is accessible to all physicists and engineers.
 Then we proceed to the case of multiple independent degrees of freedom, which requires tools from differential geometry.
-To make the treatment of the latter case more accessible, we will use a notation that is similar to the one used by physicists in general relativity.
+Here we will use a notation that is similar to the one used in general relativity.
+We will concentrate mainly on the geometrical and physical meaning of the mathematical objects and their relationships, leaving the full argument and calculations in the appendix.
 

### 2022-07-20 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
ActionPrinciple: working on the appendix

@@
 
-The first thing we need to do is recast the equations of motion in a form that will make apparent both their geometrical interpretation and their physical meaning.
-The setting is phase space extended with the time variable.
-That is, the space charted by position $q$, momentum $p$ and time $t$.
-It is precisely because we have exactly three variables that the standard tools of vector calculus can be used.
-
 \begin{figure}
@@
 
-Once we have recast the equations of motion on the extended phase space, we can see how the action principle comes from a straightforward application of vector calculus in the case of divergence-free fields.
+As we want to characterize the evolution of states over time, the appropriate setting is phase space extended with the time variable.
+That is, the space charted by position $q$, momentum $p$ and time $t$ as can be seen in Fig.
+\ref{extended_phase_space}.
+In the same way that we write $x^i = \{ x, y, z \}$ for the three dimension of space, we write
+\begin{equation}\label{sdof_variables}
+	\xi^a = \{ q, p, t\}
+\end{equation}
+for the three dimensions of the extended phase space.
 
-\subsection{Hamiltonian reformulation}
-
-Let us quickly review the equations for a Hamiltonian system with a single degree of freedom in the extended phase space.
-We can chart all states at all times by using three variables, $\{q,p,t\}$.
-Given a set of initial conditions $\{q_0,p_0,t_0\}$, we can find a single evolution $\gamma = \{q(t), p(t), t\}$ by solving Hamilton's equations:
-\begin{equation}\label{Hamilton_equations}
+Under the assumption that
+\begin{align}\label{assum_detrev}
+	\tag*{(DR)}
+	\parbox{2.8in}{the system undergoes deterministic and reversible evolution}
+\end{align}
+we can define a displacement vector field
+\begin{equation}\label{sdof_displacement}
 \begin{aligned}
-	\frac{dq}{dt} &= \frac{\partial H}{\partial p} \\
-	\frac{dp}{dt} &= -\frac{\partial H}{\partial q}.
+	\vec{S} &= \left\{ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right\} \\
+	&= S^a e_a = \frac{d\xi^a}{dt} e_a .
 \end{aligned}
 \end{equation}
-How can we better understand the physics and geometry captured by these equations?
+that tells how states move in time.\footnote{Where possible, we will be writing the same expression in both vector calculus and component notations.} The time component of the displacement vector field is constrained as we have
+\begin{equation}\label{sdof_time_constraint}
+	S^t=\frac{dt}{dt}=1.
+\end{equation}
 
-Let us step back, and forget we have a Hamiltonian system. Suppose we simply have a system described by a single degree of freedom and assume the following:
-\begin{align}\label{Detrev_assumptions}
-	\parbox{2.8in}{the system undergoes deterministic and reversible evolution.}
-\end{align}
+If assumption \ref{assum_detrev} is valid, we expect the flow of states through a closed surface to be zero: as many states flow in and out of the region.
+Alternatively, if we assign a probability, or probability density, to each trajectory, the assumptions requires that probability to not change, so integrating the probability over a closed surface must yield zero.
+However we see it, assumption \ref{assum_detrev} means the field is divergence-free,\footnote{Given that this is a three dimensional space, we can use the standard tools of vector calculus.} that is
+\begin{equation}\label{sdof_div_free}
+	\nabla \cdot \vec{S} = \partial_a S^a = 0.
+\end{equation}
 
-Under this assumption, given a state at a moment in time, there is a well-defined change in all variables which is described by the displacement vector field
-\begin{equation}
-	\vec{S} = \left\{ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right\}.
-\end{equation}
-If we consider a finite region in the extended phase space, given that the evolution is deterministic and reversible we expect the number of states flowing in and flowing out of the region to be the same.
-Equivalently, we expect the evolutions to neither converge or diverge.
-Mathematically, the displacement field must be divergence-free, that is
-\begin{equation}
-	\nabla \cdot \vec{S} = 0,
-\end{equation}
-to properly represent a deterministic and reversible evolution.
-
-Since $\vec{S}$ is divergence-free, we can find a vector potential
-\begin{equation}
-	\vec{\theta} = \{\theta_q, \theta_p, \theta_t\}
-\end{equation}
-such that
-\begin{equation}\label{displacement_has_potential}
-	\vec{S} = - \nabla \times \vec{\theta}.
+Since the displacement field is divergence-free, it admits a vector potential. We have
+\begin{equation}\label{sdof_displacement_potential}
+\begin{aligned}
+	\vec{\theta} &= \{\theta_q, \theta_p, \theta_t\} \\
+	&= \theta_a e^a \\
+	\vec{S} &= - \nabla \times \vec{\theta} \\
+	&= - \epsilon^{abc} \partial_b \theta_c \, e_a. \\
+\end{aligned}
 \end{equation}
@@
 
-Given that $\vec{\theta}$ is a vector potential, we have the usual gauge arbitrariness since $\nabla \times(\vec{\theta} + \nabla f) = \nabla \times \vec{\theta}$ and therefore the displacement field $\vec{S}$ remains unchanged.
-By choosing $f(q,p,t)$ appropriately, we can set
-\begin{equation}
+Because the displacement field must satisfy \ref{sdof_time_constraint}, without loss of generality we can set
+\begin{equation}\label{sdof_potential_expression}
+\begin{aligned}
+	\vec{\theta} &= \{p, 0, -H\} \\
+	&= p e^q - H e^t.
+\end{aligned}
+\end{equation}
+By applying definition \ref{sdof_displacement} and expanding \ref{sdof_displacement_potential} with \ref{sdof_potential_expression}, we have
+\begin{equation}\label{sdof_Ham_eq}
+	\left\{ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right\} = - \nabla \times \vec{\theta} = \left\{ \frac{\partial H}{\partial p},-\frac{\partial H}{\partial q}, 1 \right\},
+\end{equation}
+which yields Hamilton's equations. Note that the arguments works in reverse: any Hamiltonian system with one degree of freedom yields a divergence-free displacement field.
+
+\begin{figure}
+	a \includegraphics[width = 0.45\textwidth]{ActionNonOptimized.png} \\
+	b \includegraphics[width = 0.45\textwidth]{ActionOptimized.png}
+	\caption{\footnotesize{The variation of the action is the flow of the displacement field $\vec{S}$ through the surface $\Sigma$ that sits between the path $\gamma$ and its variation $\gamma'$.
+In the second picture we see that the flow is zero if the path is an actual evolution of the system, since the displacement field will be parallel to the path $\gamma$ and therefore the surface $\Sigma$.}}\label{fig_action}
+\end{figure}
+
+Now that we are recast Hamiltonian mechanics using vector calculus, let us start constructing the principle of stationary action.
+As illustrated in Fig.
+\ref{fig_action}a, take a path $\gamma$ with endpoints $A$ and $B$, not necessarily a solution of the equations of motion.
+Then take a variation $\gamma'$ of that path and identify a surface $\Sigma$ between them.
+We can ask: what is the flow of the displacement field $\vec{S}$ through $\Sigma$?
+Because $\vec{S}$ is divergence-free, the flow through $\Sigma$ will depend only on the contour, therefore the question is well posed.
+We find
+\begin{equation}\label{sdof_action}
+\begin{aligned}
+	- \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} &= \iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
+	=  \oint_{\partial \Sigma} \vec{\theta}  \cdot d\vec{\gamma} 
+	&= \delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}.
+\end{aligned}
+\end{equation}
+
+Now suppose $\gamma$ is a solution of the equation of motion, as in Fig.
+\ref{fig_action}b.
+Then $\gamma$ is a field line and the flow is tangent to $\Sigma$ no matter what $\gamma'$ we picked.
+The converse is true: if we look for those paths for which the flow through $\Sigma$ is zero no matter what $\gamma'$, we find a solution to the equation of motion.
+The solutions, then, are those paths and only those paths for which
+\begin{equation}\label{sdof_stationary_action}
+	\begin{aligned}
+		0 &=\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} \\
+		&= \delta \int^{t_2}_{t_1} \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt} dt \\  
+		&= \delta \int^{t_2}_{t_1} p \frac{dq}{dt} - H dt .		
+	\end{aligned}
+\end{equation}
+
+Lastly we need to be able to express conjugate momentum in terms of the velocity. This can be done if we assume that 
+\begin{align}\label{assum_kineq}
+	\tag*{(KE)}
+	\parbox{2.8in}{the kinematic of the system is enough to reconstruct its dynamics.}
+\end{align}
+This means that by looking at just the trajectory in space $q(t)$, we are able to reconstruct the state at each moment in time.
+Therefore we must be able to write $p=p(q,\dot{q})$, and therefore we can also write
+\begin{equation}\label{sdof_Lagrangian}
+	L(q, \dot{q}) = p(q, \dot{q}) \frac{dq}{dt} - H(q, p(q, \dot{q})).
+\end{equation}
+
+We find that a system for which \ref{assum_detrev} and \ref{assum_kineq} are valid can be characterized in terms of the action principle with a suitable Lagrangian.
+The converse is also true: if the action principle allows for a unique solution, then a Hamiltonian can be written and therefore the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
+
+We have thus demistified the action principle, and turned it into a geometrical property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field.
+We also have a clear physical meaning: the principle of stationary action is equivalent to assuming determinism/reversibility \ref{assum_detrev} and kinematic equivalence \ref{assum_kineq}.
+However, we do feel that the principle expresses these requirements in a very roundabout way.
+
+\section{Multiple degrees of freedom}
+
+To generalize to multiple degree of freedom we have to abandon the tools provided by vector calculus and embrace the ones provided by differential geometry.
+For $N$ independent degrees of freedom, we will have the $2N+1$ variables
+\begin{equation}\label{mdof_variables}
+	\xi^a = \{ q^i, p_i, t\}.
+\end{equation}
+The displacement vector field will be
+\begin{equation}\label{mdof_displacement}
+	\begin{aligned}
+		\vec{S} &= S^a e_a = \frac{d\xi^a}{dt} e_a =\frac{dq^i}{dt} e_{q^i} + \frac{dp_i}{dt} e_{p_i} + \frac{dt}{dt} e_t
+	\end{aligned}
+\end{equation}
+which also satisfies \ref{sdof_time_constraint}.
+
+The flow of $\vec{S}$ will still be divergence-free, but this property only tells us that the total number of states is conserved. We also assumed that
+\begin{align}\label{assum_indep}
+	\tag*{(IND)}
+	\parbox{2.8in}{the degrees of freedom are independent.}
+\end{align}
+This means that, at each moment in time, the total number of states (i.e.
+the volume) of a parallelopided will be the product of states identified by each degree of freedom (i.e.
+the areas).
+To be able to quantify the number of states identified by each degree of freedom, we introduce the rank-2 tensor (more precisely a two-form)
+\begin{equation}\label{mdof_form}
+	\omega = \omega_{ab} \, e^a \otimes e^b,
+\end{equation}
+which we call the state counting form.\footnote{Note that this approach is consistent with statistical mechanics.}
+
+The counting form must be of rank two as independent degree of freedom are bi-dimensional (i.e.
+quantity plus conjugate).
+Given two vectors $\vec{v}$ and $\vec{w}$, these identify a parallelogram in phase space and
+\begin{equation}\label{mdof_form_applied}
+	\omega(\vec{v}, \vec{w}) = \omega_{ab} v^a w^b
+\end{equation}
+quantifies the number of states over its surface.
+
+The form $\omega$ will need to be anti-symmetric
+\begin{equation}\label{mdof_form_antisymm}
+	\omega(\vec{v}, \vec{w}) = - \omega(\vec{w}, \vec{v}).
+\end{equation}
+as the parallelogram identified by $\vec{v}$ and $\vec{w}$ in that order will be the same as the one identified by $\vec{w}$ and $\vec{v}$ with opposite orientation.
+The form $\omega$ will also be closed, meaning\footnote{In our notation, $d\Sigma$ represents the infinitesimal surface element of integration, which is the argument of the form.}
+\begin{equation}\label{mdof_closed_form}
+	\oiint_\Sigma \omega(d\Sigma) = 0
+\end{equation}
+over all closed contractible surfaces.
+This stems from \ref{assum_indep}.
+Suppose you take a surface and translate it along another independent variable.
+The count of states changed along the translation cannot depend on the independent variable.
+Therefore if we construct any parallelepiped at equal time, two opposite sides will contain the same number of states with opposite orientation, and therefore the integral of $\omega$ over the whole surface would be zero.
+
+Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative\footnote{We write the exterior derivative as $\partial \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.} of a covector $\theta$:
+\begin{equation}\label{mdof_form_potential}
+	\begin{aligned}
+		\theta &= \theta_a e^a = \theta_{q^i} e^{q^i} + \theta_{p_i} e^{p_i} + \theta_t e^t \\
+		\omega &= - \partial \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
+	\end{aligned}
+\end{equation}
+The minus sign is the convention in symplectic geometry.
+
+The above is the generalization of \ref{sdof_displacement_potential} and it is important to note the differences.
+The exterior derivative takes the place of the curl and it is $\omega$, not $\vec{S}$, that can be expressed in terms of the potential $\theta$.
+Yet, there is an important geometrical relationship we still need to take into account: the direction of the displacement does not contribute new states.
+Physically, this is a consequence of \ref{assum_detrev}: states cannot be created or disappear over time.
+There we have
+\begin{align}\label{mdof_displacement_kills}
+	\omega(\vec{S}, \cdot) = 0.
+\end{align}
+Mathematically we say that the displacement vector kills the form, it makes it zero no matter what vector we put on the other side.
+
+Similarly to the single d.o.f., without loss of generalize we can set
+\begin{equation}\label{mdof_potential_expression}
+\begin{aligned}
+	\theta &= \{p_i, 0, -H\} \\
+	&= p_i e^{q^i} - H e^t.
+\end{aligned}
+\end{equation}
+and calculate the components of the counting form
+\begin{equation}\label{mdof_form_components}
+	\omega_{ab} = (\partial\theta)_{ab} = \begin{bmatrix}
+		0 & \delta^j_i & - \partial_{q^i} H \\
+		-\delta^i_j & 0 & - \partial_{p_i} H \\
+		\partial_{q^j} H & \partial_{p_j} H & 0
+	\end{bmatrix}.
+\end{equation}
+
+By combining \ref{mdof_displacement}, \ref{mdof_displacement_kills} and \ref{mdof_form_components} one finds
+\begin{equation}\label{mdof_Ham_eq}
+\begin{aligned}
+	S^{p_j} &= \partial_{q^j} H \\
+	S^{q^j} &= - \partial_{p_j} H
+\end{aligned}
+\end{equation}
+which recovers Hamilton's equations.
+
+The setup to recover the action principle is the same, except that we need to substitute vector calculus with differential geometry.
+First we find that the flow through $\Sigma$ is the variation of the action using Stokes' theorem
+\begin{equation}\label{mdof_action}
+	\begin{aligned}
+		- \iint_{\Sigma} \omega(d\Sigma) &= \delta \int_{\gamma} \theta(d\gamma) \\
+		&= \delta \int^{t_2}_{t_1} p_i \frac{dq_i}{dt} - H dt .
+	\end{aligned}
+\end{equation}
+If $\gamma$ is a solution of the equations of motions, then $\frac{d\gamma}{dt} = \vec{S}$. By \ref{mdof_displacement_kills}, $\vec{S}$ is the only vector field that kill the form $\omega$ therefore
+\begin{equation}\label{mdof_stationary_action}
+	\begin{aligned}
+		- \iint_{\Sigma} \omega(d\Sigma) &= \iint_{\Sigma} \omega(\vec{S}, d\lambda) dt = 0
+	\end{aligned}
+\end{equation}
+only for the solutions. This recovers the action principle for multiple degrees of freedoms.
+
+\section{Discussion and conclusions}
+
+We have found that the principle of stationary action is equivalent to assumptions \ref{assum_detrev}, \ref{assum_indep} and \ref{assum_kineq} and that the variation of the action can be understood as the flow of evolutions through the surface delimited by the path and its variation.
+This allows us to fully understand both the physical and geometrical significance of the principle in a much more precise way.
+
+Note that the true physical content is in the displacement field $\vec{S}$ and the counting form $\omega$ whose properties descend from the assumption.
+Both the Lagrangian and the potential $\theta$ are not uniquely defined and are subject to a convenient, though still arbitrary, choice of gauge.
+The Lagrangian and the action, then, do not directly encode physical property of the system.
+Saying that ``nature chooses the path that minimizes the action'', therefore, is not really that insightful because such a quantity is not even uniquely defined.
+On the other hand, the three equivalent assumptions give us a much more clear picture.
+
+Another interesting element is that a version of the principle of stationary actions holds even if \ref{assum_kineq} does not apply.
+In this case the ``Lagrangian'' $L(q,\dot{q},p)$ depends on conjugate momentum as well.
+Also note that the cases where the Lagrangian does not admit a Hamiltonian are exactly the cases where the action principle fails to yield a unique solution.
+In this sense, the action principle is better understood as a feature of Hamiltonian mechanics, instead of Lagrangian mechanics.
+
+As a final comment, note that privileging the Hamiltonian picture is more in line with quantum mechanics; the idea of $\omega$ as a tool to count states is in line with statistical mechanics; the expression \ref{mdof_potential_expression} is remarkably similar to that of the relativistic four-momentum.
+This is one of the key insights we get from our Reverse Physics approach: there is a unity among the different physical theories that begs to be brought to light.
+We are convinced that a version of this geometrical understanding must exist in the quantum world.
+
+\bibliography{bibliography}
+
+\section*{Appendix}
+
+In this section we will present the arguments in full, including some important mathematical details.
+Given that this article is still aimed at a physics audience, rather than on mathematical rigor, will be on a precise map between physical concepts and their mathematical representation.
+We divide the appendix into three section: the first will address the single d.o.f.
+case, the second will give a geometric interpretation of the tools of differential geometry that is suitable for physicists, and lastly we address the multiple d.o.f.
+case.
+
+\subsection*{One degree of freedom}
+
+For the case of one degree of freedom we are going to use vector calculus.
+We can because the problem can be studied with three variables (position, momentum and time), which is the only setting where vector calculus truly works.
+In this sense, nature really gave us a hand as three dimensional spaces are much more intuitive for us, therefore we can develop the ideas in a familiar context before generalizing.
+
+We break down the problem into sequential steps, which serves as a summary for the whole argument. We will have to:
+\begin{enumerate}[label=(\roman*)]
+	\item show that assuming determinism and reversibility \ref{assum_detrev} is equivalent to assuming that the displacement field $\vec{S}$ defined in \ref{sdof_displacement} exists and is divergence-free as in \ref{sdof_div_free}
+	\item show that the vector potential $\vec{\theta}$ for $\vec{S}$ defined in \ref{sdof_displacement_potential} can be expressed, without loss of generality, as \ref{sdof_potential_expression}
+	\item show that \ref{sdof_div_free} and \ref{sdof_potential_expression} recover Hamilton's equations as in \ref{sdof_Ham_eq}
+	\item show that the flow of $\vec{S}$ through a surface $\Sigma$ delimited by a path $\gamma$ and its variation $\gamma'$ is the variation of the integral of the vector potential, as claimed in \ref{sdof_action}
+	\item show that the action is stationary for and only for the solutions of the equations of motion
+	\item show that the action can be expressed as functions of position, velocity and time if and only if \ref{assum_kineq} applies.
+\end{enumerate}
+
+For (i), determinism and reversibility means that giving the state at some point in time means constraining the state at all times, that the system can evolve in only one way.
+In math terms, if $X$ is the manifold of all possible states at times, given $P = \{q_0, p_0, t_0\}$ there will be only one possible evolution $\gamma : \mathbb{R} \to X$ such that $\gamma(t_0) = P$.
+
+If $X$ were in a space, the previous mathematical condition would be sufficient, but on a manifold it is not.
+Consider a damped harmonic oscillator: it satisfies the previous condition, yet the system will concentrate at the equilibrium.
+What needs to happen is that densities also need to be mapped.
+That is, if a certain fraction of the system is in a given initial state, the same fraction of the system must be in the corresponding final state.
+Equivalently, if the system has a given probability of starting in a given initial state, it must have the same probability of ending in the corresponding final state.
+In mathematical terms, if $\rho : X \to \mathbb{R}$ is a statistical distribution (or a probability density) defined on our extended space time and $\gamma$ is a possible evolution, we must have $\rho(\gamma(t_0)) = \rho(\gamma(t_1))$.
+
+This requirement tells us two things.
+First, is that $X$ must be a differentiable manifold.
+To define densities over a set of variables and map them back and forward in time Jacobians must be defined, which means the derivative across state variables must be defined as well, in particular with respect with time, which plays the double role of a coordinate of $X$ and the evolution parameter.
+This means that the evolutions must be differentiable curves and therefore $\vec{S}$ must be defined which leads to \ref{sdof_displacement}.
+The double role of time gives us \ref{sdof_time_constraint}.
+
+Secondly, since the densities must be constant over evolutions, the flow of these densities through a closed surface in $X$ must be zero: each evolution will enter and exit the region delimited by the surface, giving no net contribution.
+Mathematically this requires the displacement to be divergence-free, which leads to \ref{sdof_div_free}.
+
+For (ii), since $\vec{S}$ is divergence-free, we can find a vector potential as shown in \ref{sdof_displacement_potential}.
+Note that the vector potential is not uniquely identified, since $\nabla \times(\vec{\theta} + \nabla f) = \nabla \times \vec{\theta}$ and therefore the displacement field $\vec{S}$ remains unchanged.
+This is what is commonly referred to as gauge freedom in physics, and it is analogous to what happens for a magnetic field or for an incompressible fluid.
+
+We use this gauge freedom to set the momentum component to zero.
+In fact, we can always choose $f(q,p,t)$ such that $\partial_p f = -\theta_p$ and redefine $\vec{\theta}$ to be $\vec{\theta} + \nabla f$.
+Therefore we can set
+\begin{equation*}
 	\vec{\theta} = \{\theta_q, 0, \theta_t\}
-\end{equation}
+\end{equation*}
 without loss of generality.
 
-The displacement field has a further constraint, since $S_t = dt/ dt = 1$. Given \ref{displacement_has_potential}, we have
+We now use the constraint \ref{sdof_time_constraint} on the time component. We have
 \begin{align*}
-	S_t &= - \left(\frac{\partial}{\partial q}  \theta_p - \frac{\partial}{\partial p}  \theta_q\right) \\
+	S_t = \frac{dt}{dt} &= - \left(\frac{\partial}{\partial q}  \theta_p - \frac{\partial}{\partial p}  \theta_q\right) \\
 	&= - \left(\frac{\partial}{\partial q}  0 - \frac{\partial}{\partial p}  \theta_q\right) \\
@@
 Therefore we have:
-\begin{equation}
+\begin{equation*}
 	\vec{\theta} = \{p, 0, \theta_t\}
-\end{equation}
+\end{equation*}
 Lastly, we rename the last component as $\theta_t = -H$ which leads to 
-\begin{equation}\label{potential_final}
-	\vec{\theta} = \{p, 0, -H\}.
-\end{equation}
+\begin{equation*}
+	\vec{\theta} = \{p, 0, -H\}
+\end{equation*}
+and recovers \ref{sdof_potential_expression}.
 
-If we now expand each component of \ref{displacement_has_potential} using \ref{potential_final}, we have:\footnote{Note that $\frac{dp}{dt}$ is not the same as $\frac{\partial p}{\partial t}$.
-In the first case we are taking a total derivative along the evolution, and therefore the momentum can change.
-In the second case we are taking a partial derivative at constant $q$ and $p$ by definition, and therefore $\frac{\partial p}{\partial t}=0$.}
+For (iii), we expand each component of \ref{sdof_displacement_potential} using \ref{sdof_potential_expression}.
+Again note the double role of time: $\frac{dp}{dt}$ is not the same as $\frac{\partial p}{\partial t}$.\footnote{We could be more precise and use $\tau$ for the evolution parameter, leaving $t$ just for the coordinate.
+The treatment would automatically acquire a relativistic flavor and would enlighten us to other interesting results, though it would distract from the main purpose of this paper, the action principle.
+} In the first case we are taking a total derivative along the evolution, and therefore the momentum can change.
+In the second case we are taking a partial derivative at constant $q$ and $p$ by definition, and therefore $\frac{\partial p}{\partial t}=0$.
+We have
 \begin{align*}
@@
 \end{align*}
-We have recovered Hamilton's equations \ref{Hamilton_equations} as the equations for a deterministic and reversible system.
+We have recovered Hamilton's equations \ref{sdof_Ham_eq} as the equations for a deterministic and reversible system.
 The only thing that the equations say is that states move in time at the same rate (i.e.
@@
 
-\subsection{Action principle}
+Note that if we started from Hamilton's equation, the system would be deterministic and reversible: position and momentum at a fixed time would indeed allow us to pick a single differentiable evolution and, by Liouville's theorem, densities would be constant along such evolution.
+Therefore we have an equivalence between assumption $\ref{assum_detrev}$ and Hamilton's equations $\ref{sdof_Ham_eq}$ in the single degree of freedom case.
 
-Let us now review the principle of stationary action.
-We chart all states at all times by using $\{q, \frac{dq}{dt}, t\}$.
-We have a Langrangian $L(q, \frac{dq}{dt}, t)$ and, given a path $\gamma = \{q(t), \frac{dq}{dt}(t), t\}$ with $t_1 \leq t \leq t_2$, we define the action as
-\begin{equation}
-	\mathcal{A}[\gamma] = \int_{t_1}^{t_2} L(q, \frac{dq}{dt}, t) dt.
-\end{equation}  An actual evolution of the system is such that the action is stationary:
-\begin{equation}
-	\frac{\delta \mathcal{A}}{\delta \gamma} = 0.
-\end{equation}
-The connection between the Lagrangian and Hamiltonian formulation is given by the Legendre transformation
-\begin{equation}
-	L = p \frac{dq}{dt} - H.
-\end{equation}
-How can we connect this with the previous picture?
+Now that we have developed a clear geometrical and physical understanding of Hamilton's equations, we can concentrate on the action principle.
+For (iv) we can apply Stokes' theorem.
+We take a path $\gamma$ and a variation $\gamma'$.
+Since they share the endpoint, taken together they form a closed line where $\gamma'$ is taken with the opposite orientation.
+Because $\vec{S}$ is divergence-free, the flow of $\vec{S}$ through a surface depends only on its contour.
+Moreover, by Stoke's theorem, the flow through the surface will be equal to the integral of the potential over the contour.
+Putting it all together we have:
+\begin{align*}
+	- \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} &=
+	\iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
+	&= \oint_{\partial \Sigma} \vec{\theta}  \cdot d\vec{\gamma} \\
+	&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma'} \\
+	&= \delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}.
+\end{align*}
+This recovers \ref{sdof_action}.
 
-Let us step back, forget about Lagrangians, and resume the discussion from before.
-As shown in fig.
-\ref{action}-a, let $\gamma$ be a path in the extended phase space with endpoints $A$ and $B$, not necessarily a solution of the equations of motion.
-That is, $d\vec{\gamma}$ will not be, in general, equal to $\vec{S}dt$.
-Consider the line integral of the vector potential $\vec{\theta}$ along $\gamma$.
-We have:
+For (v), note that in the neighborhood of $\gamma$ the infinitesimal surface element $d\vec{Sigma}$ has one side along $\gamma$ itself.
+Therefore we can write $d\vec{Sigma} = d\vec{\gamma} \times d\vec{\lambda}$, where $d\vec{\lambda}$ is the infinitesimal displacement that connects $\gamma$ with $\gamma'$.
+The integrand then becomes the triple product
+\begin{equation*}
+	\vec{S} \cdot d\vec{\gamma} \times d\vec{\lambda}.
+\end{equation*}
+We can understand it geometrically as the volume element given by the displacement $\vec{S}$, the infinitesimal line element $d\vec{\gamma}$ and the infinitesimal variation $d\vec{\lambda}$ that moves $\gamma$ to $\gamma'$.
+
+If $\gamma$ is an actual solution of the equation of motion, $\gamma$ is a field line of $\vec{S}$: $\gamma$ is tangent to $\vec{S}$ at all points.
+We can write $d\vec{\gamma} = \vec{S} dt$.
+The triple product, in this case, is zero regardless of what $d\vec{\lambda}$ is.
+We can reason the other way: we look for a line $\gamma$ such that the triple product is zero regardless of what $d\vec{\lambda}$ is.
+The only way for that to happen is if $\gamma$ is tangent to $\vec{S}$ at each point, which means $\gamma$ is a solution of the equation of motion.
+This means that the paths that make the line integral of the vector potential stationary are those and only those that solve the equations of motions.
+This recovers the principle of the stationary action in Hamiltonian form.
+
+We want to stress that this variational principle applies to any divergence-free field. For example, for a magnetic field we could write:
+\begin{equation*}
+	\iint_{\Sigma} \vec{B} \cdot d\vec{\Sigma} = \delta \int_{\gamma} \vec{A} \cdot d\vec{\gamma}.
+\end{equation*}
+We would find that the field lines of $\vec{B}$ are the lines and only the lines for which the integral of the vector potential is stationary.
+Similarly, we would find that the streamlines of an incompressible fluids are the lines and only the lines for which the integral of the vector potential of the flow is stationary.
+While these properties may be of limited practical use, the geometrical insight is the same.
+
+For (vi), if the assumption \ref{assum_kineq} of kinematic equivalence  holds, we can from a trajectory in space $q(t)$ reconstruct the state at any time.
+That is, if we are given $q(t)$ we can reconstruct the full evolution $\gamma(t)$ in the extended phase space.
+
+We saw before that the evolution must be differentiable, which means $q(t)$ must also be differentiable.
+Therefore the velocity $\dot{q} = \frac{dq}{dt}$ is defined at every instant.
+Using assumption \ref{assum_detrev}, we know that each evolution is fully identified by position and momentum at a given time.
+Therefore the space of all possible evolutions is two-dimensional.
+Given that for each evolution $\gamma(t)$ there is one and only one spatial trajectory $q(t)$, the space of all spatial trajectories must also be two-dimensional and therefore fully identified by position and velocity at a given instant.
+Therefore there must be a map $p(q, \dot{q}, t)$ that allows us to reconstruct the momentum at a given time by knowing position and velocity at that time.
+
+We can use this map to express the principle of stationary action only in terms of position and velocity we have
 \begin{equation}
 	\begin{aligned}
-	\int_A^B \vec{\theta} \cdot d\vec{\gamma} &= \int^{t_2}_{t_1} \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt} dt \\
-	&= \int^{t_2}_{t_1} \left(\theta_q \frac{dq}{dt} + \theta_p \frac{dp}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
-	&= \int^{t_2}_{t_1} \left(p \frac{dq}{dt} + 0 \frac{dp}{dt} - H \frac{dt}{dt}\right) dt \\
-	&= \int^{t_2}_{t_1} \left(p \frac{dq}{dt} - H\right) dt \\
-	&= \int^{t_2}_{t_1} L \, dt
+		\delta \int_\gamma \vec{\theta} \cdot d\vec{\gamma} &= \delta \int^{t_2}_{t_1} \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt} dt \\
+		&= \delta \int^{t_2}_{t_1} \left(\theta_q \frac{dq}{dt} + \theta_p \frac{dp}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
+		&= \delta\int^{t_2}_{t_1} \left(p \frac{dq}{dt} + 0 \frac{dp}{dt} - H \frac{dt}{dt}\right) dt \\
+		&= \delta \int^{t_2}_{t_1} \left(p \frac{dq}{dt} - H\right) dt \\
+		&= \delta \int^{t_2}_{t_1} L(q, \dot{q}, t) \, dt
 	\end{aligned}
 \end{equation}
-Therefore the line integral of the vector potential over a path $\gamma$ coincides with the action $A[\gamma]$.
+We recognize the principle of stationary action in Lagrangian form.
 
-\begin{figure}
-	a \includegraphics[width = 0.45\textwidth]{ActionNonOptimized.png} \\
-	b \includegraphics[width = 0.45\textwidth]{ActionOptimized.png}
-	\caption{\footnotesize{The variation of the action is the flow of the displacement field $\vec{S}$ through the surface $\Sigma$ that sits between the path $\gamma$ and its variation $\gamma'$.
-In the second picture we see that the flow is zero if the path is an actual evolution of the system, since the displacement field will be parallel to the path $\gamma$ and therefore the surface $\Sigma$.}}\label{action}
-\end{figure}
+Note that the argument can proceed in the opposite way.
+If we have a system for which the principle of stationary action yields a unique solution, then we can write $p=\frac{\partial L}{\partial \dot{q}} = p(q, \dot{q}, t)$ which means we are able to reconstruct the momentum, and therefore the state, at each time.
+The system satisfies \ref{assum_kineq}.
+Moreover, we can find a unique Hamiltonian such that the solution of the equations of motion are exactly those paths that satisfy the principle of stationary actions.
+Therefore the system satisfies \ref{assum_detrev} as well.
 
-Now let us consider the variation of the line integral $\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}$ caused by an infinitesimal change in the path.
-The original path $\gamma$ together with its variation $\gamma'$ form a contour $\partial \Sigma$ which encloses the surface $\Sigma$.
-We can therefore use Stokes' theorem to transform the line integral of $\vec{\theta}$ over $\partial \Sigma$ to a surface integral of the curl $\nabla \times \vec{\theta}$ over $\Sigma$.
-That is:
-\begin{align*}
-	\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} 
-	&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma'} \\
-	&= \oint_{\partial \Sigma} \vec{\theta}  \cdot d\vec{\gamma} \\
-	&= \iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
-	&= - \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma}.
-\end{align*}
-Note that, precisely because $\vec{S}$ is divergence-free, it doesn't matter which particular surface $\Sigma$ we pick, as long as its contour $\partial \Sigma$ consists of $\gamma$ and $\gamma'$.
+Before proceed to the generalizations, a few observations are in order.
+\begin{itemize}
+	\item We stress that the vector potential $\vec{\theta}$ in not uniquely defined by the physics, and therefore neither are the action and the Lagrangian.
+Giving a precise physical characterizations of those two objects, then, is hard.
+In our view, it is a hopeless task.
+	\item We found that the Hamiltonian $H$ is the time component of the vector potential $\vec{\theta}$, not a scalar. This relativistic aspect emerges naturally out of the assumptions.
+	\item One can, in principle, write Lagrangian systems for which the action principle does not yield a single solution (e.g.
+for $L=0$ all paths are stationary).
+Consistently, assumption \ref{assum_detrev} fails.
+\end{itemize}
 
-The variation of the action, then, corresponds to the flow of $\vec{S}$ through $\Sigma$.
-It tells us how many actual evolutions go through the surface.
-Note that if $\gamma$ is an evolution of the system, it is always tangent to $\vec{S}$.
-This means that the surface $\Sigma$ enclosed by $\gamma$ and its infinitesimal variation is also tangent to $\vec{S}$ and therefore $\vec{S} \cdot d\vec{\Sigma} = 0$.
-Which means the actual evolutions are those paths $\gamma$ for which
-\begin{equation}
-	\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} = 0.
-\end{equation}
-This recovers the principle of stationary action.
+\subsection*{Introduction to differential topology}
 
-The action principle, then, is a direct consequence of the divergence-free nature of the displacement field $\vec{S}$.
-Every evolution $\gamma$ is a field line of $\vec{S}$.
-Any infinitesimal surface $\Sigma$ enclosed by an evolution and its variation, then, will also be tangent to the displacement field $\vec{S}$ and therefore the surface integral of $\vec{S}$ through $\Sigma$ will be zero.
-Given that $\vec{S}$ is divergence-free, the integral through $\Sigma$ can be expressed as the line integral of the vector potential $\vec{\theta}$ over the contour of $\Sigma$.
-Asking for a stationary action means asking for the flow of $\vec{S}$ to be zero across the surface given by the path and its variation, which happens only if the path is tangent to $\vec{S}$ and is therefore an evolution.
-In practice, the principle of stationary action is a very roundabout way to ask that the path be always tangent to the displacement vector field.
+The generalization to multiple d.o.f will require the use of differential topology.
+These tools are not typically taught in a standard physics curriculum.
+If they are, the treatment is the dense and abstract one coming from mathematical tradition, which we have found to be arcane even to some mathematicians: at a topology conference, a math graduate student confessed he sticked with topology because differential geometry was too abstract for him, comment that was taken with sympathy by others.
 
-\subsection{Discussion}
+Apart from the abstract definitions, the notation gets in the way of the physics.
+For example, writing $\frac{\partial}{\partial x^i}$ for the vector basis makes sense if one is studying the space of linear operators defined by derivations.
+When studying the space of velocities, however, writing $\vec{v} = v^i\frac{\partial}{\partial x^i}$ decreases understanding.
+Therefore we will use a notation and nomenclature that is closer to what physicists already use (e.g.
+Einstein notation) and need to picture.
+For example, we will use $e_i$ for vector basis and $e^i$ for covector (dual) basis.
+To those that believe we should just use the ``correct'' mathematical nomenclature, we retort that there is no such thing.
+Notations and names are chosen within a certain context to make the points of the discourse more clear.
 
-Before proceeding to the more general case, there are two important things to note.
+What we will give here, then, is not a fully rigorous introduction to the subject, but rather a geometrical understanding of what these tool describe and a physical motivation as to why they are useful.
+We hope this will make it easier to follow the next section and to build a geometrical and physical intuition of the principle of stationary actions, which is the whole point of this article.
 
-First of all, the true (and only) physical content of the discussion lies in condition \ref{Detrev_assumptions} that assumes deterministic and reversible motion.
-The displacement field $\vec{S}$ and its properties are the only mathematical entities that are directly physically meaningful.
-The vector potential $\vec{\theta}$ is not uniquely defined by condition \ref{Detrev_assumptions}, though it is extremely useful as it allows us to map the space of all possible displacement fields of interest with a single ``scalar'' function $H$, the time component of the vector potential.
+%This very result, and other connected to it, became apparent to us precisely because we started to translate the language of differential geometry into a form that is less abstract.
+These results have been in front of our known noses, hidden in plain sight by a notation, useful to some mathematicians, but arcane to others.
+While we don't claim our notation is ``the best'', it is better for our purposes than the standard ones.
+We understand that there is a cost associated to learning a new notation, but to us that cost is essentially learning exactly why the tools are useful and what they represent in physics.
+It is worth the long term investment
 
-Since the Lagrangian is equal to the scalar product between the displacement along a path $\frac{d\vec{\gamma}}{dt}$ and the vector potential $\vec{\theta}$, it suffers all the problems of the vector potential: it is not unique and its value is not directly physically meaningful.
-The same is true for the action itself.
-While the Lagrangian and the action are clearly useful tools in physics, this result warrants caution in relying exclusively on them to understand the physical relevance and meaning of a theory.
+%In our opinion, , and In order to fully understand the geometrical and physical meaning of the general case, it helps to have a crisp intuition for the tools of differential topology.
+We will therefore go through the basic motivation of differential topology from a physicist perspective, so that we can see what differential forms and exterior derivatives are capturing at a a physical level.
 
-The second point is that our geometrical and physical interpretation works for all Hamiltonians.
-Technically, not all Hamiltonian systems admit a Lagrangian, as one needs an invertible map between conjugate momentum $p$ and the velocity $\frac{dq}{dt}$.
-For example, the following Hamiltonian
-\begin{equation}
-	H = c |p^i|
-\end{equation}
-is the one for a free photon treated as a particle. This gives us
-\begin{equation}
-	\frac{dq}{dt}= c \frac{p}{|p|}
-\end{equation}
-which means the velocity allows us to recover the direction of the conjugate momentum, but not its norm (i.e.
-the trajectory of a photon is not enough to recover its full state).
-In these cases, the Lagrangian approach cannot be applied because $\vec{\theta} \cdot \frac{d\vec{\gamma}}{dt}$ cannot be expressed as a function of position and velocity.
-Yet, if we express $L(q,p,t)$ in terms of conjugate momentum instead of velocity, the modified version of the variational principle still works.
+To boil it done, the objective is to characterize those quantities that are associated with a region of space and can be understood as sums of independent contributions.
+The mass $m$ within a region $V$ can be understood as the sum of the contributions of a mass density $\rho$ within each infinitesimal volume $dV$:
+\begin{equation*}
+	m(V) = \iiint_V \rho dV.
+\end{equation*}
+The magnetic flux $\Phi$ through a surface $\Sigma$ can be understood as the sum of the contributions of the magnetic field $\vec{B}$ over each infinitesimal $d\Sigma$:
+\begin{equation*}
+	\Phi(\Sigma) = \iint_\Sigma \vec{B} \cdot \vec{d\Sigma}.
+\end{equation*}
+The work $W$ over a path $\lambda$ can be understood as the sum of the contributions of a force field $\vec{f}$ over each infinitesimal line $d\gamma$:
+\begin{equation*}
+	W(\gamma) = \int_\gamma \vec{f} \cdot \vec{d\gamma}.
+\end{equation*}
 
-Note that the failure can happen in the reverse direction too: there are Lagrangian systems for which we cannot express the Hamiltonian in terms of conjugate momentum.
-However, these are exactly the cases where the action principle fails to yield a single solution.
-Therefore we have the following points:
-\begin{enumerate}
-	\item the only Lagrangian systems in which the action principle yields a single solution are the ones that admit a Hamiltonian formulation;
-	\item all Hamiltonian systems admit an action principle formulation, even if they do not admit a Lagrangian.
+Though the dimensionality changes, a pattern emerges: the functionals $W$, $\Phi$ and $m$ all take a region of space while $\vec{f}$, $\vec{B}$ and $\rho$ act on infinitesimal regions.
+However, the vector representation gives us three different enough objects (i.e.
+vector, pseudo-vector and pseudo-scalar) and operations, and therefore fails fail to provide a nice generalization.
+
+Suppose we wrote
+\begin{equation*}
+\begin{aligned}
+	W(\gamma) &= \int_\gamma f(d\gamma) \\
+	\Phi(\Sigma) &= \iint_\Sigma B(d\Sigma) \\
+	m(V) &= \iiint_V \rho(dV).
+\end{aligned}
+\end{equation*}
+Here we don't have vectors, but functions of infinitesimal regions, which we call differential forms.
+The force, in this notation, is a one-form (or covector) as it takes one dimensional infinitesimal regions (i.e.
+vectors); the magnetic field is a two-form; the density is a three-form.
+We can also say that a scalar field, like the temperature, is a zero-form.
+
+An infinitesimal region can also be understood as a parallelopiped, which is fully identified by its sides.
+Therefore a differential form can be understood as acting on a set of vectors that matches the dimensionality of the form.
+A one-form will take one vector, a two-form two vectors and so on.
+All forms must be anti-symmetric because switching the order of the sides does not change the parallelopiped, just its orientation.
+
+This picture is much easier to generalize to any dimensionality.
+If we note $S^k$ the space of all the $k$ dimensional subregions, a $k$-functional $F : S^k \to \mathbb{R}$ is a linear functional that takes a $k$-surface and returns a number.
+This can be expressed as
+\begin{equation*}
+	F(\sigma) = \int_\sigma \omega(d\sigma)
+\end{equation*}
+the integral of a $k$-form $\omega : V^k \to \mathbb{R}$ that takes an infinitesimal $k$-surfaces (or equivalently a set of $k$ vectors) and returns a number.
+
+This allows us to identify properties and theorems that apply to all functionals and forms regardless of their dimensionality.
+For example, if $\sigma_1$ and $\sigma_2$ are two disjoint regions, $F(\sigma_1 \cup \sigma_2) = F(\sigma_1) + F(\sigma_2)$, which also tells us $F(\emptyset) = 0$ for any functional.
+Additionally, $k$-forms are linear maps and can be written in terms of the tensor product of the co-basis $e^a$.
+We have:
+\begin{equation*}
+	\begin{aligned}
+		f &= f_a e^a \\
+		B &= B_{ab} e^a \otimes e^b \\
+		\rho &= \rho_{abc} e^a \otimes e^b \otimes e^c.
+	\end{aligned}
+\end{equation*}
+These tensor are all anti-symmetric as the forms are anti-symmetric.
+
+%If $F_k$ is the set of all linear $k$-functionals and $\Omega_k$ the set of all linear $k$-forms, we can understand $\Omega_k$ as the infiniteimsal verson of $F_k$.
+
+%of one, two or three dimensions
+
+%We say that $m$ is a functional that takes a three dimensional region and returns a number.
+This functional is linear, in the sense that it respects the limit and it is additive over disjoint regions.
+That is, $m(V_1 \cup V_2) = m(V_1) + m(V_2)$ if $V_1 \cup V_2 = \emptyset$.
+
+%Since $k$-form are also linear maps, they form a vector space and can be written in terms of a basis.
+In the same way we have a basis $e_a$ for vectors, we will have a basis $e^a$ for one-forms (or covectors) such that $e^a(e_b) = \delta^a_b$.
+We have:
+%Additionally, all forms are anti-symmetric in all indexes: a changing the order of the vectors reduces to a change of the orientation.
+Mathematically, this is why one introduces the wedge product $\wedge$ which is an anti-symmetrized version of the tensor product.
+Given that physicists are already familiar with the idea of anti-symmetric tensors, we think it's better to not introduce the wedge product and work in terms of tensors.
+
+We can see how this works for a force applied along a path. The infinitesimal displacement along $\gamma$ is given by
+\begin{equation*}
+	d\gamma = dx^a e_a.
+\end{equation*}
+Therefore 
+\begin{equation*}
+	\begin{aligned}
+		W(\gamma) &= \int_\gamma f(d\gamma) = \int_\gamma  f_a e^a(dx^b e_b) \\
+		&= \int_\gamma  f_a dx^b e^a( e_b)
+		= \int_\gamma  f_a dx^b \delta^a_b
+		= \int_\gamma  f_a dx^a.
+	\end{aligned}
+\end{equation*}
+
+Of all generalized properties, the most impressive one is certainly the generalized Stokes' theorem. Compare the divergence (or Gauss's) theorem:
+\begin{equation*}
+	\iiint_V (\nabla \cdot \vec{B}) \, dV = \oiint_{\Sigma = \partial V} \vec{B} \cdot \vec{d\Sigma}
+\end{equation*}
+with Stokes' theorem
+\begin{equation*}
+	\iint_\Sigma (\nabla \times \vec{f}) \cdot \vec{d\Sigma} = \oint_{\gamma = \partial \Sigma} \vec{f} \cdot \vec{d\gamma}
+\end{equation*}
+and the divergence theorem
+\begin{equation*}
+	\int_\gamma (\nabla \varphi) \cdot \vec{d\gamma} = \left[ \varphi \right]_A^B .
+\end{equation*}
+In all three cases we have a region where we define the integral of some differential operator applied to a field.
+This is equated to the integral over the boundary of the field itself.
+But again, the vector calculus notation prevents us to generalize as the differential operator is somewhat different in the three cases.
+
+To understand conceptually that a generalization must exist, suppose we have a $k$-functional $F$ and its corresponding $k$-form $\omega$.
+By definition $F$ acts on $k$-surfaces.
+Now suppose we have a $k+1$-dimensional surface $\Sigma$.
+This is not a valid argument for $F$.
+However, its boundary $\partial \Sigma$ will be a $k$-dimensional surface, which is a valid argument for $F$.
+Therefore from any $F$ we can create a $k+1$-functional, which we can call exterior functional of $F$, noted $\partial F$, such that $\partial F(\Sigma) = F(\partial \Sigma)$.
+If we look at all three cases above, that is exactly what is happening: the gradient, the curl and the divergence lead to a functional one dimension higher than the original.
+
+Now, since every functional has a corresponding form, $\partial F$ must have a corresponding $k+1$-form which must be fully identified by $\omega$. We note it as $\partial \omega$ and write
+\begin{equation*}
+	\partial F(\Sigma) = \int_\Sigma \partial \omega(d\Sigma) = \oint_{\sigma = \partial \Sigma} \omega(d\sigma) = F(\partial \Sigma).
+\end{equation*}
+This is the generalized Stokes' theorem and $\partial \omega$ is the exterior derivative. We use $\partial$ as we leave $d$ reserved for the infinitesimal regions.
+
+With a few steps, one can see that the divergence, curl and gradient correspond to the exterior derivative in three dimensions for zero-form, one-form and two-form respectively, and that the three theorems of vector calculus are particular instances of this generalized one.
+
+The identities
+\begin{equation*}
+	\begin{aligned}
+		\nabla \cdot \nabla \times \vec \vec{f} &= 0 \\
+		\nabla \times \nabla \varphi &= 0
+	\end{aligned}
+\end{equation*}
+both correspond to the single identity
+\begin{equation*}
+	\partial \partial \omega = 0,
+\end{equation*}
+which states that the exterior derivative applied twice is also zero.
+This identity can be understood in terms of the functionals.
+We have and $\partial \partial F(\Sigma) = F(\partial \partial \Sigma)$ and the boundary of a boundary $\partial \partial \Sigma = \emptyset$ is always the empty set.
+So, no matter what $F$ or $\Sigma$ are, we are evaluating a functional on an empty set, which is zero.
+
+In the same way that a divergence-free field can be written as the curl of a vector potential, or a curl-free field can be written as the gradient of a scalar potential, a form whose exterior derivative $\partial \omega = 0$ is zero (a closed form), can be written as the exterior derivative $\omega = \partial \theta$ of a lower dimensional form, at least is some neighborhood.
+In terms of functionals, this corresponds to the case where $F(\sigma)=0$ for all contractible surfaces, those surfaces that can be morphed to a point with a continuous transformation.
+If $F(\sigma)=0$ over all surfaces, instead, then the correspond is exact.
+
+The overall point here is that understanding differential topology in terms of functionals of regions gives us much better motivation and intuition for the math.
+
+\iffalse
+
+%If we compute the exterior derivative for zero, one and two-form
+
+%The last element we need to introduce is the exterior derivative.
+We are actually much better equipped to understand this operation as it is much more intuitive on the linear functionals than the corresponding forms.
+
+%First, note that every $k$-surface has a $k-1$ boundary.
+For example, a three dimensional volume would be enclosed by a two dimensional surface.
+Therefore we have a boundary operator $\partial \sigma^k \in S^{k-1}$.
+If the surface does not have a boundary, like the surface of a sphere, then the boundary is the empty set $\emptyset$.
+Note that the all boundaries are closed surfaces.
+Therefore the boundary of a boundary is the empty set.
+That is $\partial \partial \sigma = 0$ for all $\sigma$.
+
+Now recall Stokes' theorem:
+On the left side we have a two-functional, a functional of a surface.
+On the right side we have a one-functional, a functional of a line.
+The two-functional can be defined in terms of the one-functional: given a surface, take its boundary and pass it to the one functional.
+Recall Gauss's theorem:
+Something similar is happening except we have a three functional on the left and a two functional on the right. Recall the gradient theorem:
+\begin{equation*}
+	\int_\gamma \nabla \varphi \cdot d\gamma = \left[ \varphi \right]_A^B
+\end{equation*}
+Here we have a one functional on the left and function of points on the right, which can be understood as a zero functional.
+We note a pattern, but again the vector calculus notation does not help us express the pattern.
+
+Differential topology generalizes the pattern through the generalized Stokes' theorem
+\begin{equation*}
+	\int_\sigma \partial \omega(d\sigma) = \oint_{\partial \sigma} \omega(d\partial\sigma)
+\end{equation*}
+Our notation differs from the standard in that we use $\partial$ as the symbol for the exterior derivative and we make explicit the argument of the forms.
+We still have, however, to understand what is it that we are generalizing.
+This is better understood in terms of linear functionals.
+
+Suppose we have a $k$-functional $F$, for example work along a path.
+Each $k$ dimensional surface $\sigma$ will have a boundary $\partial \sigma$ that is one dimensional.
+In the example, a two dimensional surface will identify a line.
+This means that, to each $k$-dimensional surface $\sigma$ we can associate a number $F(\partial \sigma)$.
+In the example, to a two-dimensional surface we can associate the work along its boundary.
+Proceeding this way we can define the exterior functional, which we note $\partial F$ and note that $\partial F(\sigma) = F(\partial \sigma)$.
+The notation is admittedly a bit unusual, but it allows us to show very easily that the exterior functional of the exterior functional is always zero.
+In fact
+\begin{equation*}
+	\partial \partial F(\sigma) = \partial F(\partial \sigma) = F(\partial \partial \sigma) = F(\emptyset) = 0
+\end{equation*}
+This is a direct consequence of the boundaries not having boundaries.
+
+Since every $k$-functional has a corresponding $k$-form, so will the exterior functionals.
+Using Stokes' theorem we find that the form associated to the exterior functional is exactly the exterior derivative.
+
+Obviously a full treatment of all these concepts goes beyond the scope of this appendix.
+Yet, we believe it is important to see that we can have a good understanding of these tools in terms of ideas that can be meaningful to physicists.
+
+\fi
+
+\subsection*{Multiple degrees of freedom}
+
+The case of multiple degrees of freedom proceeds in the same way as the single degree of freedom. We will have to:
+\begin{enumerate}[label=(\roman*)]
+	\item show that assuming determinism and reversibility \ref{assum_detrev} and independence of degrees of freedom \ref{assum_indep} is equivalent to assuming the existence of the displacement field $\vec{S}$ defined by \ref{mdof_displacement} and the closed counting form $\omega$ defined by \ref{mdof_form} such that $\omega$ is killed only by the displacement field $\vec{S}$ as in \ref{mdof_displacement_kills} or by its rescaling $\varphi \vec{S}$.
+	\item show that the potential $\theta$ for $\omega$ defined in \ref{mdof_form_potential} can be expressed, without loss of generality, as \ref{mdof_potential_expression} and that $\omega$ can be expressed as $\ref{mdof_form_components}$
+	\item show that \ref{mdof_form_potential}, \ref{mdof_displacement_kills} and \ref{mdof_potential_expression} recover Hamilton's equations as in \ref{mdof_Ham_eq}
+	\item show that the integral of $\omega$ through a surface $\Sigma$ delimited by a path $\gamma$ and its variation $\gamma'$ is the variation of the integral of the vector potential, as claimed in \ref{mdof_action}
+	\item show that the action is stationary for and only for the solutions of the equations of motion
+	\item show that the action can be expressed as functions of position, velocity and time if and only if \ref{assum_kineq} applies.
 \end{enumerate}
 
-In our view, this is enough evidence to conclude that the Hamiltonian setting in the extended phase space is the most natural to understand the action principle.
+For (i), \ref{assum_detrev}  will lead to the existence of a displacement field $\vec{S}$ that tells us how state move in the extended phase space.
+As we saw in the previous case, this is not sufficient to impose the full condition: we need to be able to quantify the number of states as they flow in time.
+To this end, we need a functional $\mathcal{F}$ that takes an hyper-surfaces $\Sigma$ in the extended phase space and returns the count of states.
+We have
+\begin{equation}
+	\mathcal{F}(\Sigma) = \int_\Sigma \Omega(d\Sigma)
+\end{equation}
+where $\Omega$ is the related form.
+Given the one-to-one correspondence between states and evolutions under \ref{assum_detrev}, $\Omega$ can be understood as quantifying either the flow of states through the surface or the states that are present on the surface.
 
-\section{Multiple degrees of freedom}
+As in the previous case, assumption \ref{assum_detrev} tells us that the integral over a closed hyper-surface is zero, since as many states flow in and out. We have
+\begin{equation}
+	\begin{aligned}
+		\oint_{\Sigma = \partial V} \Omega(d\Sigma) &= 0 = \int_V \partial \Omega(dV) \\
+		\partial\Omega &= 0.
+	\end{aligned}
+\end{equation}
+which tells us that $\Omega$ is closed.
 
-All our previous points will hold true in the generalization to multiple degrees of freedom.
-The only additional key insight is that determinism and reversibility will apply to each independent degree of freedom, meaning that we do not only preserve the total number of states of the system, but also that the total number of states is the product of the configurations given by each independent degree of freedom.
+Given that $\Omega$ quantifies the flow over $\vec{S}$, the flow for any surface that is tangent to $\vec{S}$ must vanish. If $d\gamma = \vec{S} \varphi dt$ for some scalar $\varphi$, we must have
+\begin{equation}
+	\Omega(d\gamma, \cdot) = \varphi dt \Omega(\vec{S}, \cdot) = 0.
+\end{equation}
+The notation is a bit improper as $\vec{S}$ is not technically an infinitesimal displacement, but a rate of displacement given a parameter (i.e.
+a velocity).
+However, the slight abuse of notation makes the concept more clear: the direction of motion does not contribute to the flow.
 
-To express this requirement, the tools provided by differential calculus are not enough.
-We need elements of differential geometry, particularly symplectic geometry and contact manifolds.
-To make the treatment more appropriate for a physics audience, we are going to use a notation that parallels Einstein notation for general relativity.
-\footnote{The main differences are as follows.
-The vector basis will be noted as $e_a$ instead of $\frac{\partial}{\partial x^a}$.
-The covector basis will be noted as $e^a$ instead of $dx^a$.
-The exterior derivative will be noted as $\partial$ (e.g.
-$\partial \omega$) instead of $d$ (e.g.
-$d \omega$).}
+On the other hand, this must be the only direction that does not contribute to the flow.
+Since $\vec{S}$ cannot be zero because states must flow in time, for any other direction $\vec{T}$ that is not parallel to $\vec{S}$ there must be a surface tangent to $\vec{T}$ and not $\vec{S}$.
+The flow through that surface cannot be zero.
+Therefore the only way that the flow is zero for any surface tangent to a vector $\vec{T}$ is if $\vec{T}= \varphi \vec{S}$
 
-\subsection{Hamiltonian reformulation}
+We now turn to assumption \ref{assum_indep}.
+This requires the count of states to factorize: if we have a parallelopiped in phase space, the count of states (i.e.
+its volume) is the product of the count of the configurations over each initial conditions (i.e.
+the area of the sides).
+Equivalently, \ref{assum_indep} must allow the case of a distribution over states with no correlations between the independent d.o.f., in which cases the density must factorize.
+Since we must be able to quantify the configuration for each d.o.f., and since each d.o.f.
+is two dimensional, we must have a two-form $\omega$ such that:
+\begin{equation}
+	\Omega = \omega^N.
+\end{equation}
+
+Given that $\Omega$ is closed, we have
+\begin{equation}
+	\begin{aligned}
+		\partial \Omega &= \partial \omega \wedge \omega^{N-1} = 0 \\
+		\partial \omega &= 0
+	\end{aligned}
+\end{equation}
+which means that $\omega$ is also closed.
+This result can be also argued on physical grounds.
+If we imagine to translate a surface along an independent variable, the number of configurations identified by the surface cannot change.
+Therefore, if we imagine a parallelopiped, opposite sides must identify the same number of configurations.
+The integral will take the sides with an opposite orientation, and therefore there will be zero net contribution.
+
+Putting it all together, assuming \ref{assum_detrev} and \ref{assum_indep} gives us the existence of the displacement vector field $\vec{S}$ and the closed counting two-form $\omega$ such that $\vec{S}$ identifies the only vector field that kills $\omega$.
+
+The converse is also true: if we start with the mathematical requirement we recover the physical assumptions.
+If the displacement vector field $\vec{S}$ is the only vector field that kills the two-form $\omega$, then $\Omega = \omega^N$ is non-zero and is closed form.
+The system is deterministic and reversible because the displacement is well defined and it is paired of a flow that conserves the count of states.
+The degrees of freedom are independent because $\Omega = \omega^N$, but it can be seen in a more direct way.
+
+If each pair of state variables $q^i$, $p_i$ forms an independent degree of freedom, the only surfaces spanned by matching position and momentum will define actual configurations, while all other combinations will not define any.
+Therefore:
+\begin{equation}\label{canonical_conditions}
+	\begin{aligned}
+		\omega(e^{q^i}, e^{p_j}) = \omega_{q^i p_j} &= \delta^i_j = - \omega_{p_j q^i} \\
+		\omega(e^{q^i}, e^{q^j}) = \omega_{q^i q^j} &= 0 \\
+		\omega(e^{p_i}, e^{p_j}) = \omega_{p_i p_j} &= 0
+	\end{aligned}
+\end{equation}
+If we start with a closed form $\omega$ that has a single degenerate direction, by Darboux theorem we can find coordinates for the non-degenerate part $\omega$ that satisfy the above conditions.
 
@@
 As degrees of freedom are bi-dimensional, this means quantifying the areas of two-dimensional surfaces.
-Therefore we introduce the rank-2 tensor
+Therefore we introduce the rank-2 tensor (more precisely a two-form)
 \begin{equation}

### 2022-07-20 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
ActionPrinciple: more work on the appendix

@@
 
-The converse is also true: if we start with the mathematical requirement we recover the physical assumptions.
-If the displacement vector field $\vec{S}$ is the only vector field that kills the two-form $\omega$, then $\Omega = \omega^N$ is non-zero and is closed form.
-The system is deterministic and reversible because the displacement is well defined and it is paired of a flow that conserves the count of states.
-The degrees of freedom are independent because $\Omega = \omega^N$, but it can be seen in a more direct way.
+The converse is also true: if we start with the mathematical requirements, we can recover the physical assumptions.
+If the displacement vector field $\vec{S}$ is the only vector field that kills the two-form $\omega$, then $\Omega = \omega^N$ is a non-zero closed form.
+The displacement is well defined and it is paired of a flow that conserves the count of states, therefore the system is deterministic and reversible.
+The degrees of freedom are independent because $\Omega = \omega^N$, and we can recover a suitable set of independent state variables.
 
@@
 
-Our system will be now composed of $N$ independent degrees of freedom.
-This means that we can chart all states at all times using $2N + 1$ variables $\xi^a = \{ q^i, p_i, t\}$.
-We will use $\xi^a$ when we want to span all variables; $q^i$ and $p_i$ when we want to span the position or the momentum of all degrees of freedom.
-
-As before, we will have a displacement field
-\begin{equation}
-	\vec{S} = \frac{d\xi^a}{dt} e_{a} = \frac{dq^i}{dt} e_{q^i} + \frac{dp_i}{dt} e_{p_i} + \frac{dt}{dt} e_t.
-\end{equation}
-However, unlike in the single degree of freedom case, $\vec{S}$ is not the object used fully characterize the geometry of the space.
-
-The main idea is that we want to be able to quantify the number of states identified by each degree of freedom.
-As degrees of freedom are bi-dimensional, this means quantifying the areas of two-dimensional surfaces.
-Therefore we introduce the rank-2 tensor (more precisely a two-form)
-\begin{equation}
-	\omega = \omega_{ab} \, e^a \otimes e^b,
-\end{equation}
-which we call the state counting tensor. The idea is that a pair of vectors $\vec{v}$ and $\vec{w}$ will identify a parallelogram in phase space and 
-\begin{equation}
-	\omega(\vec{v}, \vec{w}) = \omega_{ab} v^a w^b
-\end{equation}
-quantifies the number of states over its surface.
-
-The tensor $\omega$ will need to satisfy additional conditions to be physically meaningful.
-First of all, it will need to be anti-symmetric.
-The parallelogram identified by $\vec{v}$ and $\vec{w}$ in that order will be the same as the one identified by $\vec{w}$ and $\vec{v}$ therefore the area is the same.
-However, the switch changes handedness, and therefore will introduce a minus sign.
-Therefore we have:
-\begin{equation}
-	\omega(\vec{v}, \vec{w}) = - \omega(\vec{w}, \vec{v}).
-\end{equation}
-Since the rank-2 tensor is anti-symmetric, it can be understood as a two-form in differential geometry.
-Only the components below the diagonal are independent, which means $\sum_1^{2N}i =N(2N+1)$ independent components.
-
-The form $\omega$ will also be closed, meaning
-\begin{equation}\label{mdof_closed_form}
-	\oiint \omega = 0
-\end{equation}
-over all closed surfaces.
-In fact, consider a parallelepiped.
-Given that the degrees of freedom are independent, and we have deterministic and reversible motion, two opposite sides will contain the same number of states.
-Given that the two opposite sides will have opposite orientation, their total contribution will be zero.
-This logic can be extended to all three-dimensional differential volumes, as they can be decomposed into infinitesimal parallelepiped.
-
-Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative of a covector $\theta$. We have
-\begin{equation}
-\begin{aligned}
-	\theta &= \theta_a e^a = \theta_{q^i} e^{q^i} + \theta_{p_i} e^{p_i} + \theta_t e^t \\
-	\omega &= - \partial \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
-\end{aligned}
-\end{equation}
-The minus sign is the convention in symplectic geometry.
-
-In the generalization we need to distinguish between vectors, like $\vec{S}$, covectors, like $\theta$, and two-forms, like $\omega$.
-The single degree of freedom used only three variables, therefore to each parallelogram in $\{q, p, t\}$ there is only one perpendicular direction, which allows us to blur the line between the objects in the following way:
-\begin{itemize}
-	\item for each covector $\theta$ we can find a vector $\vec{\theta}$ such that $\theta(\vec{S})=\vec{\theta} \cdot \vec{S}$;
-	\item for each two-form $\omega$ we can find a vector $\vec{\omega}$ such that $\omega(\vec{v}, \vec{w}) = \vec{\omega} \cdot \vec{v} \times \vec{w}$.
-\end{itemize}
-This is what allows us to think only in terms of vectors.
-This cannot be done in the general case.
-In particular, the generalization of the curl is no longer an operation between vectors, but an operation that takes a one-form and returns a two-form.
-
-Now we need to further characterize $\omega$.
-Each pair of variables $q^i$, $p_i$ forms an independent degree of freedom.
-This means that only surfaces spanned by matching position and momentum will define actual states, while all other combinations will not define any state.
-Therefore:
-\begin{equation}\label{canonical_conditions}
-\begin{aligned}
-	\omega(e^{q^i}, e^{p_j}) = \omega_{q^i p_j} &= \delta^i_j = - \omega_{p_j q^i} \\
-	\omega(e^{q^i}, e^{q^j}) = \omega_{q^i q^j} &= 0 \\
-	\omega(e^{p_i}, e^{p_j}) = \omega_{p_i p_j} &= 0
-\end{aligned}
-\end{equation}
-Of the $N(2N+1)$ independent components given the anti-symmetry requirement, the first condition sets $N^2$ of these while the last two set $N(N-1)/2$ independent components each.
-This totals $N(2N-1) = \sum_1^{2N-1}i$ independent components.
-
-Assuming determinism and reversibility, time evolution contributes no new states, it just moves them in time.
-That is, no states appear or disappear from any degree of freedom.
-The displacement field doesn't define any state, no matter which other vector it is paired with.
-Mathematically, we say that the displacement kills the form
-\begin{align}\label{mdof_displacement_kills}
-	\omega(\vec{S}, \cdot) = 0.
-\end{align}
-This sets another $2N$ components (since $\omega(\vec{S}, \vec{S}) = 0$ is already set by anti-symmetry). All independent components are now set.
+Note that we have broken down exactly how each assumptions contributes to the mathematical setting.
+If \ref{assum_indep} fails, then $\Omega$ cannot be understood as $\omega^N$, which means we have much fewer condition on the motion: we only know that the total flow is conserved, but we cannot break it down along each d.o.f..
+On the other hand, if \ref{assum_detrev} we may not have a single displacement field $\vec{S}$ and/or the form that counts states would not be closed.
+This type of exact mapping between assumptions and their mathematical expression is one of the key goals of Reverse Physics in particular and our overall project Assumptions of Physics more in general.
+
+For (ii)
 

### 2022-07-25 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
Finished first draft of the article

@@
 \begin{equation}\label{mdof_form_components}
-	\omega_{ab} = (\partial\theta)_{ab} = \begin{bmatrix}
-		0 & \delta^j_i & - \partial_{q^i} H \\
-		-\delta^i_j & 0 & - \partial_{p_i} H \\
-		\partial_{q^j} H & \partial_{p_j} H & 0
+	\omega_{ab} = (-\partial\theta)_{ab} = \begin{bmatrix}
+		0 & \delta^j_i & \partial_{q^i} H \\
+		-\delta^i_j & 0 & \partial_{p_i} H \\
+		-\partial_{q^j} H & -\partial_{p_j} H & 0
 	\end{bmatrix}.
@@
 \begin{aligned}
-	S^{p_j} &= \partial_{q^j} H \\
-	S^{q^j} &= - \partial_{p_j} H
+	S^{q^j} &= \partial_{p_j} H \\
+	S^{p_j} &= - \partial_{q^j} H
 \end{aligned}
@@
 For the case of one degree of freedom we are going to use vector calculus.
-We can because the problem can be studied with three variables (position, momentum and time), which is the only setting where vector calculus truly works.
-In this sense, nature really gave us a hand as three dimensional spaces are much more intuitive for us, therefore we can develop the ideas in a familiar context before generalizing.
+This is possible since the problem can be studied with three variables (position, momentum and time), which is the only setting where vector calculus truly works.
+In this sense, nature gave us a hand as three dimensional spaces are much more intuitive for us, therefore we can develop the ideas in a familiar context before generalizing.
 
@@
 \begin{align*}
-	S_t = \frac{dt}{dt} &= - \left(\frac{\partial}{\partial q}  \theta_p - \frac{\partial}{\partial p}  \theta_q\right) \\
+	S^t = \frac{dt}{dt} &= - \left(\frac{\partial}{\partial q}  \theta_p - \frac{\partial}{\partial p}  \theta_q\right) \\
 	&= - \left(\frac{\partial}{\partial q}  0 - \frac{\partial}{\partial p}  \theta_q\right) \\
@@
 \begin{align*}
-	S_q = \frac{dq}{dt}
+	S^q = \frac{dq}{dt}
 	&= - \left( \frac{\partial}{\partial p} \theta_t - \frac{\partial}{\partial t} \theta_p \right) \\
@@
 \begin{align*}
-	S_p = \frac{dp}{dt}
+	S^p = \frac{dp}{dt}
 	&= - \left( \frac{\partial}{\partial t} \theta_q - \frac{\partial}{\partial q} \theta_t \right) \\
@@
 \begin{align*}
-	S_t = \frac{dt}{dt}
+	S^t = \frac{dt}{dt}
 	&= - \left( \frac{\partial}{\partial q} \theta_p - \frac{\partial}{\partial p} \theta_q \right) \\
@@
 
-Given that $\Omega$ is closed, we have
+Using assumptions \ref{assum_detrev} and \ref{assum_indep} again we can see how $\omega$ must be closed
 \begin{equation}
-	\begin{aligned}
-		\partial \Omega &= \partial \omega \wedge \omega^{N-1} = 0 \\
-		\partial \omega &= 0
-	\end{aligned}
+	\partial \omega = 0.
 \end{equation}
-which means that $\omega$ is also closed.
-This result can be also argued on physical grounds.
-If we imagine to translate a surface along an independent variable, the number of configurations identified by the surface cannot change.
+If we imagine translating a surface along an independent variable, the number of configurations identified by the surface cannot change.
+This uses \ref{assum_indep}.
+If we imagine translating in time, again the number of configurations cannot change because of \ref{assum_detrev}.
 Therefore, if we imagine a parallelopiped, opposite sides must identify the same number of configurations.
 The integral will take the sides with an opposite orientation, and therefore there will be zero net contribution.
+It is interesting to note that this condition requires both assumptions.
 
@@
 The converse is also true: if we start with the mathematical requirements, we can recover the physical assumptions.
-If the displacement vector field $\vec{S}$ is the only vector field that kills the two-form $\omega$, then $\Omega = \omega^N$ is a non-zero closed form.
-The displacement is well defined and it is paired of a flow that conserves the count of states, therefore the system is deterministic and reversible.
-The degrees of freedom are independent because $\Omega = \omega^N$, and we can recover a suitable set of independent state variables.
+If we are given a two-form $\omega$ on a $2N+1$ dimensional space, we can define the $2N$ form $\Omega = \omega^N$.
+This form will be non-zero because $\vec{S}$ is the only vector field that kills the two-form $\omega$.
+It will also be closed since
+\begin{equation}
+	\partial \Omega = \partial \omega \wedge \omega^{N-1} = 0.
+\end{equation}
+So the flow of the well defined displacement field $\vec{S}$, as measured by $\Omega$, is zero through closed surfaces. This implies \ref{assum_detrev}.
 
-If each pair of state variables $q^i$, $p_i$ forms an independent degree of freedom, the only surfaces spanned by matching position and momentum will define actual configurations, while all other combinations will not define any.
+To obtain \ref{assum_indep}, if each pair of state variables $q^i$, $p_i$ forms an independent degree of freedom, only the surfaces spanned by matching position and momentum will define actual configurations, while all other combinations will not define any.
 Therefore:
@@
 If we start with a closed form $\omega$ that has a single degenerate direction, by Darboux theorem we can find coordinates for the non-degenerate part $\omega$ that satisfy the above conditions.
+This recovers \ref{assum_indep}.
 
@@
 
-For (ii)
-
-Those familiar with symplectic and contact geometry will recognize that conditions \ref{canonical_conditions} mean that $q^i$ and $p_i$ are canonical coordinates and, by Darboux's theorem, we will be able to write the covector potential as
-\begin{equation}\label{mdof_potential_expression}
+For (ii), we note that since the form is closed, we can find a covector potential $\theta$ such that $\omega=\partial \theta$.
+Those familiar with symplectic and contact geometry will know that using Darboux's theorem, we can find a set of coordinates such that we can express $\theta$ as
+\begin{equation}
+	\tag*{(\ref{mdof_potential_expression})}
 	\theta = p_i e^{q^i} - H e^t.
 \end{equation}
-Let us convince ourselves of this result by constructing a proof that is similar to what we have done for the single degree of freedom.
+Just stating this is, of course, not very insightful.
+To gain more insight, we provide the sketch of a proof that, as we have done for the single degree of freedom, uses gauge transformations to fix the form of the potential.
 
-First, we express \ref{canonical_conditions} in terms of the covector potential. We have:
+As a starting point, we take conditions \ref{canonical_conditions} that state mathematically that $q^i$ and $p_i$ form independent degrees of freedom.
+We re-express those conditions in terms of the potential $\theta$
 \begin{equation}\label{canonical_potential_conditions}
@@
 
-At this point, we calculate $\partial\theta(S, \cdot ) $. The components of $\partial\theta$ are:
-\begin{equation*}
-	(\partial\theta)_{ab} = \begin{bmatrix}
-		0 & \delta^j_i & - \partial_{q^i} H \\
-		-\delta^i_j & 0 & - \partial_{p_i} H \\
-		\partial_{q^j} H & \partial_{p_j} H & 0
-	\end{bmatrix}
-\end{equation*}
-We have
-\begin{align*}
-	\partial\theta(S, \cdot )  &= S^a (\partial\theta)_{ab} e^b = 0 \\
-	&= (S^{q^i}(\partial\theta)_{q^ib} + S^{p_i}(\partial\theta)_{p_ib} + S^{t}(\partial\theta)_{tb}) e^b \\
-	&= (S^{q^i}(\partial\theta)_{q^iq^j} + S^{p_i}(\partial\theta)_{p_iq^j} + S^{t}(\partial\theta)_{tq^j}) e^{q^j} + \\
-	& (S^{q^i}(\partial\theta)_{q^ip_j} +  S^{p_i}(\partial\theta)_{p_ip_j} + S^{t}(\partial\theta)_{tp_j}) e^{p_j} + \\
-	& (S^{q^i}(\partial\theta)_{q^it} + S^{p_i}(\partial\theta)_{p_it} + S^{t}(\partial\theta)_{tt}) e^t \\
-	&= (-S^{p_i}\delta^i_j + S^{t}\partial_{q^j} H ) e^{q^j} + \\
-	& (S^{q^i}\delta^j_i +  S^{t}\partial_{p_j} H) e^{p_j} + \\
-	& (-S^{q^i} \partial_{q^i} H - S^{p_i} \partial_{p_i} H) e^t \\
-\end{align*}
+The insight is that the independence of the degree of freedom leads to the independent expression of the components of the covector potential.
+We stress again that this is a choice, and that the specific expression of the components is not physically constrained.
 
-All components must be zero, therefore we have the following three equations:
-\begin{align*}
-	S^{p_j} &= S^{t} \partial_{q^j} H \\
-	S^{q^j} &= - S^{t}\partial_{p_j} H \\
-	-S^{q^i} \partial_{q^i} H - S^{p_i} \partial_{p_i} H &= S^{t}\partial_{p_i} H \partial_{q^i} H - S^{t} \partial_{q^i} H \partial_{p_i} H = 0
-\end{align*}
-Note that the last expression is not a new equation: it is identical to zero given the previous two equations. Since $S^t = 1$, we have:
-\begin{align*}
-	S^{p_j} &= \partial_{q^j} H \\
-	S^{q^j} &= - \partial_{p_j} H
-\end{align*}
-which recovers Hamilton's equations for multiple degrees of freedom.
-
-Once again, we have found that the whole physical content is included in condition \ref{Detrev_assumptions}.
-The geometry in the extended phase space is fully specified by the state counting form $\omega$.
-This does not define lengths and angles, like the metric tensor does for a more standard geometric setting.
-It defines areas of two-dimensional surfaces in terms of states, and angles between two-dimensional surfaces in terms of the degree of independence between degrees of freedom.
-What the equations say is simply that states move in time at the same rate (i.e.
-$\frac{dt}{dt} = 1$) with a deterministic and reversible flow (i.e.
-no states are created or destroyed on any independent degree of freedom).
-
-\subsection{Action principle}
-Having reformulated Hamiltonian mechanics in the extended phase space using the language of differential geometry, the geometrical interpretation of the action principle is essentially the same.
-
-The Lagrangian is equal to the potential covector applied to the displacement along a generic path. That is
+The components of $\partial\theta$ are
 \begin{equation}
 	\begin{aligned}
-L &= \theta\left(\frac{d\vec{\gamma}}{dt}\right) = \theta_a \frac{d\xi^a}{dt} \\
-&= p_i \frac{dq^i}{dt} + 0 \frac{dp_i}{dt} - H \frac{dt}{dt} \\
-&= p_i \frac{dq^i}{dt} - H.
+		(\partial\theta)_{q^i p_j} &= \partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i} = \partial_{q^i} 0 - \partial_{p_j} p_i= - \delta^i_j \\
+		(\partial\theta)_{q^i q^j} &= \partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i} = \partial_{q^i}p_j - \partial_{q^j}p_i = 0 \\
+		(\partial\theta)_{p_i p_j} &= \partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i} = \partial_{p_i} 0 - \partial_{p_j} 0 = 0 \\
+		(\partial\theta)_{q^i t} &= \partial_{q^i}\theta_{t} - \partial_{t}\theta_{q^i} = \partial_{q^i} (-H) - \partial_{t} p_i= - \partial_{q^i} H \\
+		(\partial\theta)_{p_i t} &= \partial_{p_i}\theta_{t} - \partial_{t}\theta_{p_i} = \partial_{p_i} (-H) - \partial_{t} 0= - \partial_{p_i} H \\
+	\end{aligned}
+\end{equation}
+Therefore we recover
+\begin{equation}
+	\tag*{(\ref{mdof_form_components})}
+	\omega_{ab}=(-\partial\theta)_{ab} = \begin{bmatrix}
+		0 & \delta^j_i & \partial_{q^i} H \\
+		-\delta^i_j & 0 & \partial_{p_i} H \\
+		- \partial_{q^j} H & - \partial_{p_j} H & 0
+	\end{bmatrix}.
+\end{equation}
+
+For (iii) we calculate $\omega(S, \cdot ) $. We have
+\begin{equation}
+\begin{aligned}
+	\omega(S, \cdot )  &= S^a \omega_{ab} e^b = 0 \\
+	&= (S^{q^i}\omega_{q^ib} + S^{p_i}\omega_{p_ib} + S^{t}\omega_{tb}) e^b \\
+	&= (S^{q^i}\omega_{q^iq^j} + S^{p_i}\omega_{p_iq^j} + S^{t}\omega_{tq^j}) e^{q^j} + \\
+	& (S^{q^i}\omega_{q^ip_j} +  S^{p_i}\omega_{p_ip_j} + S^{t}\omega_{tp_j}) e^{p_j} + \\
+	& (S^{q^i}\omega_{q^it} + S^{p_i}\omega_{p_it} + S^{t}\omega_{tt}) e^t \\
+	&= (-S^{p_i}\delta^i_j - S^{t}\partial_{q^j} H ) e^{q^j} + \\
+	& (S^{q^i}\delta^j_i -  S^{t}\partial_{p_j} H) e^{p_j} + \\
+	& (S^{q^i} \partial_{q^i} H + S^{p_i} \partial_{p_i} H) e^t \\
 \end{aligned}
@@
 
-As before, we will consider the variation of the line integral $\delta \int_{\gamma} \theta(d\vec{\gamma})$, which can be expressed as
+All components must be zero, therefore we have the following three equations:
 \begin{equation}
 \begin{aligned}
-		\int_A^B \theta(d\vec{\gamma}) &= \int^{t_1}_{t_0} \theta\left(\frac{d\vec{\gamma}}{dt}\right) dt \\
-		&= \int^{t_1}_{t_0} \left(p_i \frac{dq^i}{dt} - H \right) dt
+	S^{p_j} &= - S^{t} \partial_{q^j} H \\
+	S^{q^j} &= S^{t}\partial_{p_j} H \\
+	S^{q^i} \partial_{q^i} H + S^{p_i} \partial_{p_i} H &= S^{t}\partial_{p_i} H \partial_{q^i} H \\
+	& - S^{t} \partial_{q^i} H \partial_{p_i} H = 0
 \end{aligned}
 \end{equation}
-The original path $\gamma$ together with its variation $\gamma'$ form a contour $\partial \Sigma$.
-We can then use Stokes' theorem to transform the line integral of $\theta$ to a surface integral of its exterior derivative.
-That is:
-\begin{align*}
-	\delta \int_{\gamma} \theta(d\vec{\gamma}) = 
-	&= \int_{\gamma} \theta(d\vec{\gamma}) - \int_{\gamma'} \theta(d\vec{\gamma}') \\
-	&= \oint_{\partial \Sigma} \theta(d\vec{\gamma}) \\
-	&= \iint_{\Sigma} \partial \theta (d\Sigma) \\
-	&= - \iint_{\Sigma} \omega(d\Sigma).
-\end{align*}
-At each point, one side of $d\Sigma$ will be $d\vec{\gamma}$, therefore $\omega(d\Sigma) = \omega(d\vec{\gamma}, d\vec{\lambda})$ where $d\vec{\lambda}$ is the direction of the variation.
-If $\gamma$ is an actual evolution, $d\vec{\gamma} = \vec{S} dt$ and, by \ref{mdof_displacement_kills}, we have
+Note that the last expression is not a new equation: it is identical to zero given the previous two equations. Since $S^t = 1$, we have
+\begin{equation}
+	\tag*{(\ref{mdof_Ham_eq})}
+\begin{aligned}
+	S^{q^j} &= \partial_{p_j} H \\
+	S^{p_j} &= - \partial_{q^j} H 
+\end{aligned}
+\end{equation}
+which recovers Hamilton's equations for multiple degrees of freedom.
+
+The argument can be run in the other direction.
+If we have a system that obeys \ref{mdof_Ham_eq}, we can define the two-form $\omega$ using \ref{mdof_form_components}.
+We can verify that $\omega$ is closed, it is killed by $\vec{S}$ and recover all the starting points.
+Therefore we find again complete equivalence with Hamiltonian mechanics and assumptions \ref{assum_detrev} and \ref{assum_indep}.
+
+For those familiar with general relativity, recall that the metric tensor $g_{\alpha\beta}$ defines the geometry by defining lengths and angles.
+Here we something similar in the sense that the two-form $\omega_{ab}$ is a rank two tensor that defines the geometry of the space.
+The difference is that $\omega_{ab}$ does not define lengths and angles, but only areas of two-dimensional surfaces, which corresponds to the number of configuration on each independent degree of freedom.
+Mathematically, this difference is between the symmetry of $g_{\alpha\beta} = g_{\beta\alpha}$ and the anti-symmetry of $\omega_{ab} = -\omega_{ba}$.
+In this context, what the equations of motion say is that states move in time at the same rate (i.e.
+$\frac{dt}{dt} = 1$) with a deterministic and reversible flow (i.e.
+no states are created or destroyed on any independent degree of freedom).
+Nothing else.
+
+Comparing \ref{mdof_form_components} and \ref{mdof_Ham_eq} we see that the time components $\omega_{at}$ are actually the components of $\vec{S}$.
+In fact, the only non-trivial components of $\omega$ are components of $\vec{S}$.
+These two objects, then, are not simply related: they express the same information.
+
+For a single degree of freedom, recalling that $S^t = 1$, we can describe $\omega$ with the following nine components
+\begin{equation}\label{mdof_reduceto_sdof}
+	\omega_{ab}=(-\partial\theta)_{ab} = \begin{bmatrix}
+		0 & S^t & -S^p \\
+		-S^t & 0 & S^q \\
+		S^p & - S^q & 0
+	\end{bmatrix}.
+\end{equation}
+In three dimensions, using the Levi-Civita symbol $\epsilon^{abc}$ we can write
+\begin{equation}
+	S^a = \epsilon^{abc} \omega_{bc}
+\end{equation}
+which tells us that $\vec{S}$ is the pseudo-vector corresponding to the direction perpendicular to $\omega$.
+This is the same relationship that we would have, for example, for the magnetic field psuedo-vector and the spatial components of the electromagnetic field in relativity.
+If we express $\omega$ in terms of $\theta$, we have
+\begin{equation}
+	S^a = - \epsilon^{abc} \partial_b \theta_c
+\end{equation}
+which recovers \ref{sdof_displacement_potential}: the displacement field is minus the curl of $\omega$. Thus we can see exactly how the general case reduces to the single d.o.f. case.
+
+For (iv), we simply need to apply Stokes' theorem.
+The situation is identical to the one degree of freedom, except that the flow now is not given by the surface integral of $\vec{S}$, but of the form $\omega$.
+The path $\gamma$ and its variation $\gamma'$ will form a closed line where $\gamma'$ is taken with the opposite orientation.
+Because $\omega$ is closed, its surface integral depends only on its contour.
+Moreover, by Stoke's theorem, the flow through the surface will be equal to the integral of the potential over the contour.
+Putting it all together we recover \ref{mdof_action}
 \begin{equation}
 \begin{aligned}
-	\delta \int_{\gamma} \theta (d\vec{\gamma}) &= - \iint_{\Sigma} \omega(d\Sigma) \\
-	&= - \iint_{\Sigma} \omega(d\vec{\gamma}, d\vec{\lambda}) \\
-	&= - \iint_{\Sigma} \omega(\vec{S}, d\vec{\lambda}) dt \\
-	&= - \iint_{\Sigma} 0 = 0
+	- \int_{\Sigma} \omega(d\Sigma) &=
+	\int_{\Sigma} \partial \theta (d\Sigma) \\
+	&= \oint_{\partial \Sigma} \theta(d\partial\Sigma) \\
+	&= \int_{\gamma} \theta(d\gamma) - \int_{\gamma'} \theta (d\gamma') \\
+	&= \delta \int_{\gamma} \theta (d\gamma).
 \end{aligned}
@@
 
-The action principle, then, is a direct consequence of the geometry set by \ref{mdof_closed_form} and \ref{mdof_displacement_kills}, which is the generalization of the divergence-free nature of the displacement field $\vec{S}$.
-Every evolution $\gamma$ is a field line of $\vec{S}$.
-Any infinitesimal surface $d\Sigma$ enclosed by an evolution and its variation, then, will also be tangent to the displacement field $\vec{S}$ and therefore will contribute no states, $\omega(d\Sigma)$ will be zero.
-Given that $\omega$ is closed, the integral over $\Sigma$ can be expressed as the line integral of the covector potential $\theta$ over the contour of $\Sigma$.
+For (v), if the path $\gamma$ is a solution to the equation of motion, it is tangent at every point to $\vec{S}$.
+Therefore we can write $d\gamma= \vec{S} dt$.
+Each infinitesimal surface element $d\Sigma$ between the line and the variation will have $d\gamma$ as one side.
+Let us casll $d\lambda$ the other side, which is the displacement that takes the path to the variation.
+We have:
+\begin{equation}
+\begin{aligned}
+	-\int_\Sigma \omega(d\Sigma) &= -\int_\Sigma \omega(d\gamma, d\lambda) \\
+	&= -\int_\Sigma \omega(\vec{S}, d\lambda) dt = 0.
+\end{aligned}
+\end{equation}
+The converse is also true: suppose the flow through $\gamma$ and $\gamma'$ is zero no matter the choice of gamma, this means that at every point $\omega(d\gamma, \cdot) dt = 0$, which means $d\gamma$ kills $\omega$.
+But since $\vec{S}$ is the only direction that kills $\omega$ at every point, $d\gamma$ must be everywhere tangent to $\vec{S}$ and therefore $\gamma$ is a solution of the equations of motion.
 
-\subsection{Discussion}
+For (vi), the argument is exactly the same as in the single d.o.f..
+As the evolution is differentiable, the velocity $\dot{q}^i = \frac{dq^i}{dt}$ is well defined.
+Under \ref{assum_kineq} there exists an invertible map $p_i=p_i(q^j, \dot{q}^j, t)$ that allows us to reconstruct the momentum at a given time by knowing position and velocity.
+We can therefore express the principle of stationary action only in terms of position and velocity.
+We have
+\begin{equation}
+	\begin{aligned}
+		\delta \int_\gamma \theta (d\gamma) &= \delta \int^{t_2}_{t_1} \theta \left(\frac{d\gamma}{dt} dt\right) \\
+		&= \delta \int^{t_2}_{t_1} \left(\theta_{q^i} \frac{dq^i}{dt} + \theta_{p_i} \frac{dp^i}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
+		&= \delta\int^{t_2}_{t_1} \left(p_i \frac{dq^i}{dt} + 0 \frac{dp_i}{dt} - H \frac{dt}{dt}\right) dt \\
+		&= \delta \int^{t_2}_{t_1} \left(p_i \frac{dq^i}{dt} - H\right) dt \\
+		&= \delta \int^{t_2}_{t_1} L(q^i, \dot{q}^i, t) \, dt.
+	\end{aligned}
+\end{equation}
+This recovers the principle of stationary action in standard form.
 
-Apart from the technicalities needed to properly frame the problem, multiple degrees of freedom do not add much conceptually.
-All geometrical and physical content within the theory is about counting states and keeping track of how they move in time, which is done by the two-form $\omega$.
-As before, the assumption of deterministic and reversible evolution does all the work.
-
-The additional key assumption is that the degrees of freedom must be independent.
-If the degrees of freedom were not independent, then the number of states within a four-dimensional parallelepiped would not be given by the product of the count of states on the sides.
-We would need additional structure.
-If the degrees of freedom started independent but didn't remain so, the displacement field $\vec{S}$ would only preserve the count of states of volumes, not over each degree of freedom, and therefore $\omega(\vec{S}, \cdot) \neq 0$.
+Again, the argument can proceed in the opposite direction.
+If we have a system for which the principle of stationary action yields a unique solution, we can write $p_i = \frac{\partial L}{\partial \dot{q}^i} = p_i(q^j, \dot{q}^j, t)$.
+This means that we can reconstruct the momentum from the kinematics, and therefore \ref{assum_kineq} is satisfied.
+Moreover, we can define a Hamiltonian $H = p_i \dot{q}^i - L$ such that the paths that satisfy the equations of motion \ref{mdof_Ham_eq} are exactly those that satisfy the principle of stationary action.
+Therefore the system satisfies \ref{assum_detrev} and \ref{assum_indep} as well.
 
@@
 
-\section{Conclusion}
-
-We have given a geometrical interpretation of the principle of stationary action that is also fully physically motivated by the assumption of deterministic and reversible evolution.
-The key insight is that a path $\gamma$ and a variation form a closed curve which encloses a surface $\Sigma$.
-If the path is an actual evolution for the system, all states flow along the surface $\Sigma$, not through it.
-Using Stokes' theorem, the zero flow condition through the surface $\Sigma$ is transformed to a zero integral condition on its boundary (i.e.
-on the path $\gamma$ and its variation) leading to the principle of least action.
-
-We have seen that both the Lagrangian and the action depend on the potential, which means they are gauge dependent.
-This makes the numerical values of those objects not strictly physical.
-We have seen that the interpretation given also works for those Hamiltonian systems that do not have a corresponding Lagrangian, meaning that the most general expression for the action principle lies on the extended-phase-space Hamiltonian formulation of classical mechanics.
-We have seen that elements from symplectic and contact geometry are needed for the generalization to multiple degrees of freedom and that they can be formulated in a language that is less abstract and closer to physics conventions.
-
-All considered, we believe this way of looking at classical mechanics provides a more complete characterization and better insights than the usual treatments.
-
 \end{document}

### 2022-07-26 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
ActionPrinciple: reviews article text

@@
 If assumption \ref{assum_detrev} is valid, we expect the flow of states through a closed surface to be zero: as many states flow in and out of the region.
-Alternatively, if we assign a probability, or probability density, to each trajectory, the assumptions requires that probability to not change, so integrating the probability over a closed surface must yield zero.
+Alternatively, if we assign a probability, or probability density, to each trajectory, the assumptions requires that probability not to change, so integrating the probability over a closed surface must yield zero.
 However we see it, assumption \ref{assum_detrev} means the field is divergence-free,\footnote{Given that this is a three dimensional space, we can use the standard tools of vector calculus.} that is
@@
 \end{equation}
-The minus sign is introduced to match conventions in differential geometry. Mathematically, this is analogous to what is done for a magnetic field or for an incompressible fluid.
+The minus sign is introduced to match conventions. Mathematically, this is analogous to what is done for a magnetic field or for an incompressible fluid.
 
@@
 \end{equation}
-which yields Hamilton's equations. Note that the arguments works in reverse: any Hamiltonian system with one degree of freedom yields a divergence-free displacement field.
+which yields Hamilton's equations.
+Note that the arguments works in reverse: any Hamiltonian system with one degree of freedom yields a divergence-free displacement field, and therefore satisfies \ref{assum_detrev}.
 
@@
 
-Now that we are recast Hamiltonian mechanics using vector calculus, let us start constructing the principle of stationary action.
+We now turn to constructing the principle of stationary action.
 As illustrated in Fig.
@@
 Because $\vec{S}$ is divergence-free, the flow through $\Sigma$ will depend only on the contour, therefore the question is well posed.
-We find
+Using Stokes' theorem, we find
 \begin{equation}\label{sdof_action}
@@
 Then $\gamma$ is a field line and the flow is tangent to $\Sigma$ no matter what $\gamma'$ we picked.
-The converse is true: if we look for those paths for which the flow through $\Sigma$ is zero no matter what $\gamma'$, we find a solution to the equation of motion.
+The converse is true: if we look for those paths for which the flow through $\Sigma$ is zero no matter what $\gamma'$, $\gamma$ must be everywhere tangent to $\vec{S}$ so we find a solution to the equation of motion.
 The solutions, then, are those paths and only those paths for which
 \begin{equation}\label{sdof_stationary_action}
-	\begin{aligned}
-		0 &=\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} \\
-		&= \delta \int^{t_2}_{t_1} \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt} dt \\  
-		&= \delta \int^{t_2}_{t_1} p \frac{dq}{dt} - H dt .		
-	\end{aligned}
-\end{equation}
-
-Lastly we need to be able to express conjugate momentum in terms of the velocity. This can be done if we assume that 
+	0 =\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} = - \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} 
+\end{equation}
+We say this is the principle of stationary action in Hamiltonian form.
+
+The last step is to express the principle exclusively in terms of kinematic variables: position, time and velocity. This can be done if we assume that 
 \begin{align}\label{assum_kineq}
@@
 \begin{equation}\label{sdof_Lagrangian}
-	L(q, \dot{q}) = p(q, \dot{q}) \frac{dq}{dt} - H(q, p(q, \dot{q})).
+	\begin{aligned}
+		\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} 
+		&= \delta \int^{t_2}_{t_1} \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt} dt \\  
+		&= \delta \int^{t_2}_{t_1} \left(p \frac{dq}{dt} - H \right) dt \\
+		&= \delta \int^{t_2}_{t_1}L(q, \dot{q}, t) dt = 0.
+	\end{aligned}
 \end{equation}
@@
 We find that a system for which \ref{assum_detrev} and \ref{assum_kineq} are valid can be characterized in terms of the action principle with a suitable Lagrangian.
-The converse is also true: if the action principle allows for a unique solution, then a Hamiltonian can be written and therefore the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
+The converse is also true: if the action principle allows for a unique solution, then conjugate momentum and Hamiltonian are well defined and the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
 
@@
 To generalize to multiple degree of freedom we have to abandon the tools provided by vector calculus and embrace the ones provided by differential geometry.
-For $N$ independent degrees of freedom, we will have the $2N+1$ variables
+For $N$ independent degrees of freedom, we will have the $2N+1$ manifold charted by the variables
 \begin{equation}\label{mdof_variables}
@@
 
-The flow of $\vec{S}$ will still be divergence-free, but this property only tells us that the total number of states is conserved. We also assumed that
+The flow of $\vec{S}$ will still be divergence-free, but this property only tells us that the total number of states is conserved. We also need to assume that
 \begin{align}\label{assum_indep}
@@
 \end{equation}
-The minus sign is the convention in symplectic geometry.
+The minus sign is required to match conventions.
 
@@
 Physically, this is a consequence of \ref{assum_detrev}: states cannot be created or disappear over time.
-There we have
+We have
 \begin{align}\label{mdof_displacement_kills}
@@
 
-Similarly to the single d.o.f., without loss of generalize we can set
+Similarly to the single d.o.f., without loss of generality we can set
 \begin{equation}\label{mdof_potential_expression}
@@
 \begin{equation}\label{mdof_action}
-	\begin{aligned}
-		- \iint_{\Sigma} \omega(d\Sigma) &= \delta \int_{\gamma} \theta(d\gamma) \\
-		&= \delta \int^{t_2}_{t_1} p_i \frac{dq_i}{dt} - H dt .
-	\end{aligned}
-\end{equation}
-If $\gamma$ is a solution of the equations of motions, then $\frac{d\gamma}{dt} = \vec{S}$. By \ref{mdof_displacement_kills}, $\vec{S}$ is the only vector field that kill the form $\omega$ therefore
+	- \iint_{\Sigma} \omega(d\Sigma) = \delta \int_{\gamma} \theta(d\gamma) 
+\end{equation}
+If $\gamma$ is a solution of the equations of motions, then $\frac{d\gamma}{dt} = \vec{S}$.
+By \ref{mdof_displacement_kills}, $\vec{S}$ is the only vector field that kill the form $\omega$ therefore the solutions are such that
 \begin{equation}\label{mdof_stationary_action}
-	\begin{aligned}
-		- \iint_{\Sigma} \omega(d\Sigma) &= \iint_{\Sigma} \omega(\vec{S}, d\lambda) dt = 0
-	\end{aligned}
-\end{equation}
-only for the solutions. This recovers the action principle for multiple degrees of freedoms.
+		0 = \delta \int_{\gamma} \theta(d\gamma) = - \iint_{\Sigma} \omega(\vec{S}, d\lambda) dt. 
+\end{equation}
+
+Adding assumption \ref{assum_kineq}, we can express the variational principle in the usual form
+\begin{equation}\label{mdof_Lagrangian}
+	\begin{aligned}
+		\delta \int_{\gamma} \theta(d\gamma) &= \delta \int^{t_2}_{t_1} p_i \frac{dq_i}{dt} - H dt \\
+		 &= \delta \int^{t_2}_{t_1}L(q^i, \dot{q}^i, t) dt = 0.
+	\end{aligned}
+\end{equation}
 
@@
 The Lagrangian and the action, then, do not directly encode physical property of the system.
-Saying that ``nature chooses the path that minimizes the action'', therefore, is not really that insightful because such a quantity is not even uniquely defined.
+Saying that ``nature chooses the path that minimizes the action'', therefore, is not really that insightful because such a quantity is not uniquely defined.
 On the other hand, the three equivalent assumptions give us a much more clear picture.
@@
 Another interesting element is that a version of the principle of stationary actions holds even if \ref{assum_kineq} does not apply.
-In this case the ``Lagrangian'' $L(q,\dot{q},p)$ depends on conjugate momentum as well.
+In this case the ``Lagrangian'' $L(q,\dot{q},p,t)$ depends on conjugate momentum as well.
 Also note that the cases where the Lagrangian does not admit a Hamiltonian are exactly the cases where the action principle fails to yield a unique solution.

### 2022-07-26 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
ActionPrinciple: reviewed introduction

@@
 As part of our larger project Assumptions of Physics, we developed an approach, called Reverse Physics, which examines current theories to find a set of starting physical assumptions that are sufficient to rederive them.
-This gives a more physically motivated grounding of the classical theory and the proper setting to characterize both the physics and the geometry underlying the principle of stationary action.
-The assumption of determinism and reversible evolution corresponds to a diverge-free flow of the states, which preserves volumes in phase space (i.e.
-the count of states).
-The assumption of independence of the degrees of freedom means the volume reduces to the product of areas in each d.o.f., which is quantified by a closed two-form.
-The principle of stationary action arises as a feature of divergence-free fields and closed two-forms, as a general mathematical property of vector calculus and differential geometry\cite{souriau1997structure}.
-The additional assumption of equivalence between kinematics and dynamics (i.e.
-we can reconstruct the dynamical state simply by looking at the trajectory) is what allows us to express the principle in Lagrangian form.
+We have found that Lagrangian mechanics is equivalent to three assumptions: determinism/reversibility, independence of degree of freedom and kinematic/dynamics equivalence.
+This gives a clear physically motivated understanding of the classical theory which allows us to characterize both the physics and the geometry underlying the principle of stationary action.
+What we find is that this arises as a general mathematical feature of divergence-free fields (and closed two-forms), which is the appropriate tool to describe a flow that conserves the number of states.
+The assumption of equivalence between kinematics and dynamics (i.e.
+we can reconstruct the dynamical state simply by looking at the trajectory) is what allows us to express the principle in the usual form.
 The argument can proceed in the reverse direction: assuming the principle of stationary action recovers a dynamical system that exhibit those three physical assumptions.
 
-The mathematics needed to run the argument is not new, yet it is typically presented in a way and in a notation that is opaque to physicists.
-To provide a clear insight to the widest possible readership, we will first address the case of a single degree of freedom using standard vector calculus.
-This provides already all the geometrical and physical insights in a language that is accessible to all physicists and engineers.
-Then we proceed to the case of multiple independent degrees of freedom, which requires tools from differential geometry.
-Here we will use a notation that is similar to the one used in general relativity.
-We will concentrate mainly on the geometrical and physical meaning of the mathematical objects and their relationships, leaving the full argument and calculations in the appendix.
+The mathematics needed to run the argument is not new\cite{souriau1997structure}.
+The two aspects that have been sorely lacking, and that we provide, are (1) a clear geometrical interpretation of the action principle and (2) a tight connection between the math and the physics it represents.
+To make the result accessible to the widest audience, we first cover the case of a single degree of freedom using standard vector calculus.
+We then proceed to the case of multiple independent degrees of freedom using tools from differential geometry.
+The main article will present all the key point needed to follow the argument and its physical and geometrical meaning, leaving the mathematical details and calculations to the appendix.
 

### 2022-07-28 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
ActionPrinciple: another pass at the appendix

@@
 
-In this section we will present the arguments in full, including some important mathematical details.
-Given that this article is still aimed at a physics audience, rather than on mathematical rigor, will be on a precise map between physical concepts and their mathematical representation.
+In this section we will review the arguments in more detail.
+Given that this article is still aimed at a physics audience, the focus will be conceptual rigor, rather than mathematical rigor.
+What is important is to establish a precise map between physical concepts and their mathematical representation, so that it is clear what physical requirement is captured by each mathematical definition.
+
 We divide the appendix into three section: the first will address the single d.o.f.
@@
 
-For the case of one degree of freedom we are going to use vector calculus.
-This is possible since the problem can be studied with three variables (position, momentum and time), which is the only setting where vector calculus truly works.
-In this sense, nature gave us a hand as three dimensional spaces are much more intuitive for us, therefore we can develop the ideas in a familiar context before generalizing.
+For the case of one degree of freedom we are going to use vector calculus. This is possible since the problem can be studied with three variables (position, momentum and time)
+\begin{equation}
+\tag*{(\ref{sdof_variables})}
+	\xi^a = \{ q, p, t\}
+\end{equation}
+which is the only setting where vector calculus truly works.
+In this sense, nature gave us a hand as three dimensional spaces are much more intuitive for us, and we can therefore develop the ideas in a familiar context before generalizing.
 
@@
 \begin{enumerate}[label=(\roman*)]
-	\item show that assuming determinism and reversibility \ref{assum_detrev} is equivalent to assuming that the displacement field $\vec{S}$ defined in \ref{sdof_displacement} exists and is divergence-free as in \ref{sdof_div_free}
+	\item show that assuming determinism and reversibility \ref{assum_detrev} is equivalent to assuming the existence of the displacement field $\vec{S}$ defined in \ref{sdof_displacement} and its being divergence-free as in \ref{sdof_div_free}
 	\item show that the vector potential $\vec{\theta}$ for $\vec{S}$ defined in \ref{sdof_displacement_potential} can be expressed, without loss of generality, as \ref{sdof_potential_expression}
@@
 	\item show that the flow of $\vec{S}$ through a surface $\Sigma$ delimited by a path $\gamma$ and its variation $\gamma'$ is the variation of the integral of the vector potential, as claimed in \ref{sdof_action}
-	\item show that the action is stationary for and only for the solutions of the equations of motion
-	\item show that the action can be expressed as functions of position, velocity and time if and only if \ref{assum_kineq} applies.
+	\item show that the integral of the vector potential is stationary for and only for the solutions of the equations of motion
+	\item show that the integral of the vector potential reduces to the action if and only if \ref{assum_kineq} applies.
 \end{enumerate}
@@
 
-If $X$ were in a space, the previous mathematical condition would be sufficient, but on a manifold it is not.
+If $X$ were a discrete space, the previous mathematical condition would be sufficient, but on a manifold it is not.
 Consider a damped harmonic oscillator: it satisfies the previous condition, yet the system will concentrate at the equilibrium.
-What needs to happen is that densities also need to be mapped.
+What needs to happen is that densities also need to be mapped exactly.
 That is, if a certain fraction of the system is in a given initial state, the same fraction of the system must be in the corresponding final state.
@@
 To define densities over a set of variables and map them back and forward in time Jacobians must be defined, which means the derivative across state variables must be defined as well, in particular with respect with time, which plays the double role of a coordinate of $X$ and the evolution parameter.
-This means that the evolutions must be differentiable curves and therefore $\vec{S}$ must be defined which leads to \ref{sdof_displacement}.
-The double role of time gives us \ref{sdof_time_constraint}.
+This means that the evolutions must be differentiable curves and therefore $\vec{S}$ must be defined which leads to
+\begin{equation}
+\tag*{(\ref{sdof_displacement})}
+	\begin{aligned}
+		\vec{S} &= \left\{ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right\} \\
+		&= S^a e_a = \frac{d\xi^a}{dt} e_a .
+	\end{aligned}
+\end{equation}
+The double role of time gives us
+\begin{equation}
+\tag*{(\ref{sdof_time_constraint})}
+ 	S^t=\frac{dt}{dt}=1.
+ \end{equation}
 
 Secondly, since the densities must be constant over evolutions, the flow of these densities through a closed surface in $X$ must be zero: each evolution will enter and exit the region delimited by the surface, giving no net contribution.
-Mathematically this requires the displacement to be divergence-free, which leads to \ref{sdof_div_free}.
-
-For (ii), since $\vec{S}$ is divergence-free, we can find a vector potential as shown in \ref{sdof_displacement_potential}.
+Mathematically this requires the displacement to be divergence-free, which leads to
+\begin{equation}
+\tag*{(\ref{sdof_div_free})}
+	\nabla \cdot \vec{S} = \partial_a S^a = 0.
+\end{equation}
+
+For (ii), since $\vec{S}$ is divergence-free, we can find a vector potential
+\begin{equation}
+\tag*{(\ref{sdof_displacement_potential})}
+	\begin{aligned}
+		\vec{\theta} &= \{\theta_q, \theta_p, \theta_t\} \\
+		&= \theta_a e^a \\
+		\vec{S} &= - \nabla \times \vec{\theta} \\
+		&= - \epsilon^{abc} \partial_b \theta_c \, e_a. \\
+	\end{aligned}
+\end{equation}
 Note that the vector potential is not uniquely identified, since $\nabla \times(\vec{\theta} + \nabla f) = \nabla \times \vec{\theta}$ and therefore the displacement field $\vec{S}$ remains unchanged.
@@
 Therefore we can set
-\begin{equation*}
+\begin{equation}
 	\vec{\theta} = \{\theta_q, 0, \theta_t\}
-\end{equation*}
+\end{equation}
 without loss of generality.
@@
 We now use the constraint \ref{sdof_time_constraint} on the time component. We have
-\begin{align*}
+\begin{align}
 	S^t = \frac{dt}{dt} &= - \left(\frac{\partial}{\partial q}  \theta_p - \frac{\partial}{\partial p}  \theta_q\right) \\
@@
 	& = \frac{\partial \theta_q}{\partial p} = 1
-\end{align*}
+\end{align}
 Integrating, we have $\theta_q = p + g(q,t)$ where $g(q,t)$ is an arbitrary function.
@@
 Therefore we have:
-\begin{equation*}
+\begin{equation}
 	\vec{\theta} = \{p, 0, \theta_t\}
-\end{equation*}
+\end{equation}
 Lastly, we rename the last component as $\theta_t = -H$ which leads to 
-\begin{equation*}
-	\vec{\theta} = \{p, 0, -H\}
-\end{equation*}
-and recovers \ref{sdof_potential_expression}.
+\begin{equation}
+\tag*{(\ref{sdof_potential_expression})}
+	\begin{aligned}
+		\vec{\theta} &= \{p, 0, -H\} \\
+		&= p e^q - H e^t.
+	\end{aligned}
+\end{equation}
 
@@
 We have
-\begin{align*}
+\begin{equation}
+\begin{aligned}
 	S^q = \frac{dq}{dt}
@@
 	& = \frac{\partial H}{\partial p}
-\end{align*}
-\begin{align*}
+\end{aligned}
+\end{equation}
+\begin{equation}
+\begin{aligned}
 	S^p = \frac{dp}{dt}
@@
 	& = - \frac{\partial H}{\partial q}
-\end{align*}
-\begin{align*}
+\end{aligned}
+\end{equation}
+\begin{equation}
+\begin{aligned}
 	S^t = \frac{dt}{dt}
@@
 	& = 1
-\end{align*}
-We have recovered Hamilton's equations \ref{sdof_Ham_eq} as the equations for a deterministic and reversible system.
+\end{aligned}
+\end{equation}
+We have recovered Hamilton's equations
+\begin{equation}
+\tag*{(\ref{sdof_Ham_eq})}
+\begin{aligned}
+	\frac{dq}{dt} &= \frac{\partial H}{\partial p} \\
+	\frac{dp}{dt} &=-\frac{\partial H}{\partial q}
+\end{aligned}
+\end{equation}
+as the equations for a deterministic and reversible system.
 The only thing that the equations say is that states move in time at the same rate (i.e.
@@
 We take a path $\gamma$ and a variation $\gamma'$.
-Since they share the endpoint, taken together they form a closed line where $\gamma'$ is taken with the opposite orientation.
+Since they share the endpoints, taken together they form a closed line where $\gamma'$ is taken with the opposite orientation.
 Because $\vec{S}$ is divergence-free, the flow of $\vec{S}$ through a surface depends only on its contour.
 Moreover, by Stoke's theorem, the flow through the surface will be equal to the integral of the potential over the contour.
-Putting it all together we have:
-\begin{align*}
+Putting it all together we have
+\begin{equation}
+\tag*{(\ref{sdof_action})}
+\begin{aligned}
 	- \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} &=
@@
 	&= \oint_{\partial \Sigma} \vec{\theta}  \cdot d\vec{\gamma} \\
-	&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma'} \\
+	&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma}' \\
 	&= \delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}.
-\end{align*}
-This recovers \ref{sdof_action}.
-
-For (v), note that in the neighborhood of $\gamma$ the infinitesimal surface element $d\vec{Sigma}$ has one side along $\gamma$ itself.
-Therefore we can write $d\vec{Sigma} = d\vec{\gamma} \times d\vec{\lambda}$, where $d\vec{\lambda}$ is the infinitesimal displacement that connects $\gamma$ with $\gamma'$.
+\end{aligned}
+\end{equation}
+
+For (v), note that in the neighborhood of $\gamma$ the infinitesimal surface element $d\vec{\Sigma}$ has one side along $\gamma$ itself.
+Therefore we can write $d\vec{\Sigma} = d\vec{\gamma} \times d\vec{\lambda}$, where $d\vec{\lambda}$ is the infinitesimal displacement that connects $\gamma$ with $\gamma'$.
 The integrand then becomes the triple product
-\begin{equation*}
+\begin{equation}
 	\vec{S} \cdot d\vec{\gamma} \times d\vec{\lambda}.
-\end{equation*}
+\end{equation}
 We can understand it geometrically as the volume element given by the displacement $\vec{S}$, the infinitesimal line element $d\vec{\gamma}$ and the infinitesimal variation $d\vec{\lambda}$ that moves $\gamma$ to $\gamma'$.
@@
 This means that the paths that make the line integral of the vector potential stationary are those and only those that solve the equations of motions.
-This recovers the principle of the stationary action in Hamiltonian form.
+This recovers the principle of the stationary action in Hamiltonian form
+\begin{equation}
+\tag*{(\ref{sdof_stationary_action})}
+	0 =\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} = - \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} 
+\end{equation}
 
 We want to stress that this variational principle applies to any divergence-free field. For example, for a magnetic field we could write:
-\begin{equation*}
+\begin{equation}
 	\iint_{\Sigma} \vec{B} \cdot d\vec{\Sigma} = \delta \int_{\gamma} \vec{A} \cdot d\vec{\gamma}.
-\end{equation*}
+\end{equation}
 We would find that the field lines of $\vec{B}$ are the lines and only the lines for which the integral of the vector potential is stationary.
@@
 
-We can use this map to express the principle of stationary action only in terms of position and velocity we have
-\begin{equation}
+We can use this map to express the Hamiltonian form of principle of stationary action only in terms of position and velocity we have
+\begin{equation}
+\tag*{(\ref{sdof_Lagrangian})}
 	\begin{aligned}
@@
 \end{equation}
-We recognize the principle of stationary action in Lagrangian form.
+We recognize the principle of stationary action in standard form.
 
@@
 These tools are not typically taught in a standard physics curriculum.
-If they are, the treatment is the dense and abstract one coming from mathematical tradition, which we have found to be arcane even to some mathematicians: at a topology conference, a math graduate student confessed he sticked with topology because differential geometry was too abstract for him, comment that was taken with sympathy by others.
+If they are, their presentation is dense and abstract even to some mathematicians: at a topology conference, a math graduate student confessed he sticked with topology because differential geometry was too abstract for him, comment that was taken with sympathy by others.
 
@@
 
-To boil it done, the objective is to characterize those quantities that are associated with a region of space and can be understood as sums of independent contributions.
+The core objective of using differential topology in physics is to characterize those quantities that are associated with a region of space and can be understood as sums of independent contributions.
 The mass $m$ within a region $V$ can be understood as the sum of the contributions of a mass density $\rho$ within each infinitesimal volume $dV$:
@@
 An infinitesimal region can also be understood as a parallelopiped, which is fully identified by its sides.
-Therefore a differential form can be understood as acting on a set of vectors that matches the dimensionality of the form.
+Therefore a differential form can be understood as acting on a set of vectors, the sides of the parallelopiped, that matches the dimensionality of the form.
 A one-form will take one vector, a two-form two vectors and so on.
@@
 \end{equation*}
-the integral of a $k$-form $\omega : V^k \to \mathbb{R}$ that takes an infinitesimal $k$-surfaces (or equivalently a set of $k$ vectors) and returns a number.
+the integral of a $k$-form $\omega : V^k \to \mathbb{R}$ that takes an infinitesimal $k$-surface (or equivalently a set of $k$ vectors) and returns a number.
 
@@
 This is equated to the integral over the boundary of the field itself.
-But again, the vector calculus notation prevents us to generalize as the differential operator is somewhat different in the three cases.
+But again, the vector calculus notation prevents us to generalize as the differential operator is somewhat different in each of the three cases.
 
@@
 Therefore from any $F$ we can create a $k+1$-functional, which we can call exterior functional of $F$, noted $\partial F$, such that $\partial F(\Sigma) = F(\partial \Sigma)$.
-If we look at all three cases above, that is exactly what is happening: the gradient, the curl and the divergence lead to a functional one dimension higher than the original.
+If we look at all three cases above, that is exactly what is happening: the gradient, the curl and the divergence lead to a functional one dimension higher by applying the original operation on the boundary.
 
@@
 
-The identities
+Moving on, the identities
 \begin{equation*}
@@
 This identity can be understood in terms of the functionals.
-We have and $\partial \partial F(\Sigma) = F(\partial \partial \Sigma)$ and the boundary of a boundary $\partial \partial \Sigma = \emptyset$ is always the empty set.
+We have  $\partial \partial F(\Sigma) = F(\partial \partial \Sigma)$ and the boundary of a boundary $\partial \partial \Sigma = \emptyset$ is always the empty set.
 So, no matter what $F$ or $\Sigma$ are, we are evaluating a functional on an empty set, which is zero.
@@
 \begin{enumerate}[label=(\roman*)]
-	\item show that assuming determinism and reversibility \ref{assum_detrev} and independence of degrees of freedom \ref{assum_indep} is equivalent to assuming the existence of the displacement field $\vec{S}$ defined by \ref{mdof_displacement} and the closed counting form $\omega$ defined by \ref{mdof_form} such that $\omega$ is killed only by the displacement field $\vec{S}$ as in \ref{mdof_displacement_kills} or by its rescaling $\varphi \vec{S}$.
+	\item show that assuming determinism and reversibility \ref{assum_detrev} and independence of degrees of freedom \ref{assum_indep} is equivalent to assuming the existence of the displacement field $\vec{S}$ defined by \ref{mdof_displacement} and the closed counting form $\omega$ defined by \ref{mdof_form} such that $\omega$ is killed only by the displacement field $\vec{S}$ as in \ref{mdof_displacement_kills} or by a field everywhere tangent to $\vec{S}$.
 	\item show that the potential $\theta$ for $\omega$ defined in \ref{mdof_form_potential} can be expressed, without loss of generality, as \ref{mdof_potential_expression} and that $\omega$ can be expressed as $\ref{mdof_form_components}$
@@
 	\item show that the integral of $\omega$ through a surface $\Sigma$ delimited by a path $\gamma$ and its variation $\gamma'$ is the variation of the integral of the vector potential, as claimed in \ref{mdof_action}
-	\item show that the action is stationary for and only for the solutions of the equations of motion
-	\item show that the action can be expressed as functions of position, velocity and time if and only if \ref{assum_kineq} applies.
+	\item show that the integral of the vector potential is stationary for and only for the solutions of the equations of motion
+	\item show that the integral of the vector potential reduces to the action if and only if \ref{assum_kineq} applies.
 \end{enumerate}
 
-For (i), \ref{assum_detrev}  will lead to the existence of a displacement field $\vec{S}$ that tells us how state move in the extended phase space.
-As we saw in the previous case, this is not sufficient to impose the full condition: we need to be able to quantify the number of states as they flow in time.
+For (i), the arguments presented for the single d.o.f. still apply: \ref{assum_detrev} leads to the existence of a displacement field $\vec{S}$
+\begin{equation}
+\tag*{(\ref{mdof_displacement})}
+	\begin{aligned}
+		\vec{S} &= S^a e_a = \frac{d\xi^a}{dt} e_a =\frac{dq^i}{dt} e_{q^i} + \frac{dp_i}{dt} e_{p_i} + \frac{dt}{dt} e_t
+	\end{aligned}
+\end{equation}
+that tells us how states move in the extended phase space.
+We also need to reformulate state/density conservation in time, which we cannot do simply with a surface integral of $\vec{S}$.
 To this end, we need a functional $\mathcal{F}$ that takes an hyper-surfaces $\Sigma$ in the extended phase space and returns the count of states.
@@
 This form will be non-zero because $\vec{S}$ is the only vector field that kills the two-form $\omega$.
-It will also be closed since
+It will also be closed since $\partial \omega = 0$ and
 \begin{equation}
@@
 If \ref{assum_indep} fails, then $\Omega$ cannot be understood as $\omega^N$, which means we have much fewer condition on the motion: we only know that the total flow is conserved, but we cannot break it down along each d.o.f..
-On the other hand, if \ref{assum_detrev} we may not have a single displacement field $\vec{S}$ and/or the form that counts states would not be closed.
+On the other hand, if \ref{assum_detrev} fails we may not have a single displacement field $\vec{S}$ and/or the form that counts states would not be closed.
 This type of exact mapping between assumptions and their mathematical expression is one of the key goals of Reverse Physics in particular and our overall project Assumptions of Physics more in general.
 
-For (ii), we note that since the form is closed, we can find a covector potential $\theta$ such that $\omega=\partial \theta$.
+For (ii), we note that since the form is closed, we can find, at least in a contractible region, a covector potential $\theta$ such that $\omega=\partial \theta$.
 Those familiar with symplectic and contact geometry will know that using Darboux's theorem, we can find a set of coordinates such that we can express $\theta$ as

### 2022-07-29 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
ActionPrinciple: using square brackets for components

@@
 \ref{extended_phase_space}.
-In the same way that we write $x^i = \{ x, y, z \}$ for the three dimension of space, we write
+In the same way that we write $x^i = [ x, y, z ]$ for the three dimension of space, we write
 \begin{equation}\label{sdof_variables}
-	\xi^a = \{ q, p, t\}
+	\xi^a = [ q, p, t]
 \end{equation}
@@
 \begin{aligned}
-	\vec{S} &= \left\{ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right\} \\
+	\vec{S} &= \left[ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right] \\
 	&= S^a e_a = \frac{d\xi^a}{dt} e_a .
@@
 \begin{aligned}
-	\vec{\theta} &= \{\theta_q, \theta_p, \theta_t\} \\
+	\vec{\theta} &= [\theta_q, \theta_p, \theta_t] \\
 	&= \theta_a e^a \\
@@
 \begin{aligned}
-	\vec{\theta} &= \{p, 0, -H\} \\
+	\vec{\theta} &= [p, 0, -H] \\
 	&= p e^q - H e^t.
@@
 \begin{equation}\label{sdof_Ham_eq}
-	\left\{ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right\} = - \nabla \times \vec{\theta} = \left\{ \frac{\partial H}{\partial p},-\frac{\partial H}{\partial q}, 1 \right\},
+	\left[ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right] = - \nabla \times \vec{\theta} = \left[ \frac{\partial H}{\partial p},-\frac{\partial H}{\partial q}, 1 \right],
 \end{equation}
@@
 \begin{equation}\label{mdof_variables}
-	\xi^a = \{ q^i, p_i, t\}.
+	\xi^a = [ q^i, p_i, t].
 \end{equation}
@@
 \begin{aligned}
-	\theta &= \{p_i, 0, -H\} \\
+	\theta &= [p_i, 0, -H] \\
 	&= p_i e^{q^i} - H e^t.
@@
 \tag*{(\ref{sdof_variables})}
-	\xi^a = \{ q, p, t\}
+	\xi^a = [ q, p, t]
 \end{equation}
@@
 	\begin{aligned}
-		\vec{S} &= \left\{ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right\} \\
+		\vec{S} &= \left[ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right] \\
 		&= S^a e_a = \frac{d\xi^a}{dt} e_a .
@@
 	\begin{aligned}
-		\vec{\theta} &= \{\theta_q, \theta_p, \theta_t\} \\
+		\vec{\theta} &= [\theta_q, \theta_p, \theta_t] \\
 		&= \theta_a e^a \\
@@
 \begin{equation}
-	\vec{\theta} = \{\theta_q, 0, \theta_t\}
+	\vec{\theta} = [\theta_q, 0, \theta_t]
 \end{equation}
@@
 \begin{equation}
-	\vec{\theta} = \{p, 0, \theta_t\}
+	\vec{\theta} = [p, 0, \theta_t]
 \end{equation}
@@
 	\begin{aligned}
-		\vec{\theta} &= \{p, 0, -H\} \\
+		\vec{\theta} &= [p, 0, -H] \\
 		&= p e^q - H e^t.

### 2022-08-07 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
ActionPrinciple: reviewed with Christine

@@
 
-As part of our larger project Assumptions of Physics, we developed an approach, called Reverse Physics, which examines current theories to find a set of starting physical assumptions that are sufficient to rederive them.
-We have found that Lagrangian mechanics is equivalent to three assumptions: determinism/reversibility, independence of degree of freedom and kinematic/dynamics equivalence.
-This gives a clear physically motivated understanding of the classical theory which allows us to characterize both the physics and the geometry underlying the principle of stationary action.
-What we find is that this arises as a general mathematical feature of divergence-free fields (and closed two-forms), which is the appropriate tool to describe a flow that conserves the number of states.
+As part of our larger project Assumptions of Physics, we developed an approach, called Reverse Physics (TODO cite paper), which examines current theories to find a set of starting physical assumptions that are sufficient to rederive them.
+We have found that Lagrangian mechanics is equivalent to three assumptions: determinism/reversibility, independence of degrees of freedom and kinematics/dynamics equivalence (TODO cite long paper).
+This physically motivated understanding of the classical theory can be used to characterize both the physics and the geometry underlying the principle of stationary action.
+What we find is that this arises as a general mathematical feature of divergence-free fields (and closed two-forms), which are the appropriate tools to describe a flow that conserves the number of states.
 The assumption of equivalence between kinematics and dynamics (i.e.
 we can reconstruct the dynamical state simply by looking at the trajectory) is what allows us to express the principle in the usual form.
-The argument can proceed in the reverse direction: assuming the principle of stationary action recovers a dynamical system that exhibit those three physical assumptions.
+The argument can proceed in the reverse direction: assuming the principle of stationary action recovers a dynamical system that exhibits those three physical assumptions.
 
@@
 We then proceed to the case of multiple independent degrees of freedom using tools from differential geometry.
-The main article will present all the key point needed to follow the argument and its physical and geometrical meaning, leaving the mathematical details and calculations to the appendix.
+The main article will presents all the key point needed to follow the argument and its physical and geometrical meaning, leaving the mathematical details and calculations to the appendix.
 
@@
 
-As we want to characterize the evolution of states over time, the appropriate setting is phase space extended with the time variable.
-That is, the space charted by position $q$, momentum $p$ and time $t$ as can be seen in Fig.
+As we want to characterize the evolution of states over time, the appropriate setting is phase space extended with the time variable.(TODO find citations) That is, the space charted by position $q$, momentum $p$ and time $t$ as can be seen in Fig.
 \ref{extended_phase_space}.
-In the same way that we write $x^i = [ x, y, z ]$ for the three dimension of space, we write
+In the same way that we write $x^i = [ x, y, z ]$ for the three dimensions of space, we write
 \begin{equation}\label{sdof_variables}
@@
 	\tag*{(DR)}
-	\parbox{2.8in}{the system undergoes deterministic and reversible evolution}
+	\parbox{2.8in}{\emph{the system undergoes deterministic and reversible evolution}}
 \end{align}
@@
 \end{equation}
-that tells how states move in time.\footnote{Where possible, we will be writing the same expression in both vector calculus and component notations.} The time component of the displacement vector field is constrained as we have
+that describes how states move in time.\footnote{Where possible, we will be writing the same expression in both vector calculus and component notations.} The time component of the displacement vector field is constrained, as we have
 \begin{equation}\label{sdof_time_constraint}
@@
 
-If assumption \ref{assum_detrev} is valid, we expect the flow of states through a closed surface to be zero: as many states flow in and out of the region.
+If assumption \ref{assum_detrev} is valid, we expect the flow of states through a closed surface to be zero: as many states flow in as flow out of the region.
 Alternatively, if we assign a probability, or probability density, to each trajectory, the assumptions requires that probability not to change, so integrating the probability over a closed surface must yield zero.
@@
 \begin{aligned}
-	\vec{\theta} &= [\theta_q, \theta_p, \theta_t] \\
-	&= \theta_a e^a \\
-	\vec{S} &= - \nabla \times \vec{\theta} \\
-	&= - \epsilon^{abc} \partial_b \theta_c \, e_a. \\
+	\vec{\theta} &= [\theta_q, \theta_p, \theta_t] = \theta_a e^a \\
+	\vec{S} &= - \nabla \times \vec{\theta} = - \epsilon^{abc} \partial_b \theta_c \, e_a. \\
 \end{aligned}
@@
 \begin{aligned}
-	\vec{\theta} &= [p, 0, -H] \\
-	&= p e^q - H e^t.
-\end{aligned}
-\end{equation}
-By applying definition \ref{sdof_displacement} and expanding \ref{sdof_displacement_potential} with \ref{sdof_potential_expression}, we have
+	\vec{\theta} &= [p, 0, -H(q,p,t)] \\
+	&= p e^q - H(q,p,t) e^t,
+\end{aligned}
+\end{equation}
+where $H$ is a suitable function of $q$, $p$ and $t$. By applying definition \ref{sdof_displacement} and expanding \ref{sdof_displacement_potential} with \ref{sdof_potential_expression}, we have
 \begin{equation}\label{sdof_Ham_eq}
@@
 which yields Hamilton's equations.
-Note that the arguments works in reverse: any Hamiltonian system with one degree of freedom yields a divergence-free displacement field, and therefore satisfies \ref{assum_detrev}.
+Note that the argument works in reverse: any Hamiltonian system with one degree of freedom yields a divergence-free displacement field, and therefore satisfies \ref{assum_detrev}.
 
@@
 	\caption{\footnotesize{The variation of the action is the flow of the displacement field $\vec{S}$ through the surface $\Sigma$ that sits between the path $\gamma$ and its variation $\gamma'$.
-In the second picture we see that the flow is zero if the path is an actual evolution of the system, since the displacement field will be parallel to the path $\gamma$ and therefore the surface $\Sigma$.}}\label{fig_action}
+In (b) we see that the flow is zero if the path is an actual evolution of the system, since the displacement field will be parallel to the path $\gamma$ and therefore tangent to the surface $\Sigma$.}}\label{fig_action}
 \end{figure}
@@
 	- \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} &= \iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
-	=  \oint_{\partial \Sigma} \vec{\theta}  \cdot d\vec{\gamma} 
+	=  \oint_{\partial \Sigma = \gamma \cup \gamma'} \vec{\theta}  \cdot d\vec{\gamma} 
 	&= \delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}.
@@
 \end{equation}
-We say this is the principle of stationary action in Hamiltonian form.
+We call this the principle of stationary action in Hamiltonian form.
 
@@
 	\tag*{(KE)}
-	\parbox{2.8in}{the kinematic of the system is enough to reconstruct its dynamics.}
+	\parbox{2.8in}{\emph{the kinematics of the system is enough to reconstruct its dynamics}.}
 \end{align}
@@
 We find that a system for which \ref{assum_detrev} and \ref{assum_kineq} are valid can be characterized in terms of the action principle with a suitable Lagrangian.
-The converse is also true: if the action principle allows for a unique solution, then conjugate momentum and Hamiltonian are well defined and the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
-
-We have thus demistified the action principle, and turned it into a geometrical property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field.
+The converse is also true: if the action principle allows for a unique solution, then the conjugate momentum and the Hamiltonian are well defined and the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
+
+We have thus demystified the action principle, and turned it into a geometrical property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field in phase space.
 We also have a clear physical meaning: the principle of stationary action is equivalent to assuming determinism/reversibility \ref{assum_detrev} and kinematic equivalence \ref{assum_kineq}.
@@
 
-To generalize to multiple degree of freedom we have to abandon the tools provided by vector calculus and embrace the ones provided by differential geometry.
+To generalize to multiple degrees of freedom we have to abandon the tools provided by vector calculus and embrace the ones provided by differential geometry.
 For $N$ independent degrees of freedom, we will have the $2N+1$ manifold charted by the variables
@@
 	\tag*{(IND)}
-	\parbox{2.8in}{the degrees of freedom are independent.}
+	\parbox{2.8in}{\emph{the degrees of freedom are independent.}}
 \end{align}
 This means that, at each moment in time, the total number of states (i.e.
-the volume) of a parallelopided will be the product of states identified by each degree of freedom (i.e.
+the volume) of a parallelepiped will be the product of states identified by each degree of freedom (i.e.
 the areas).
@@
 \end{equation}
-which we call the state counting form.\footnote{Note that this approach is consistent with statistical mechanics.}
-
-The counting form must be of rank two as independent degree of freedom are bi-dimensional (i.e.
+which we call the state counting form.\footnote{Note that this approach is consistent with statistical mechanics. (TODO: add references)}
+
+The counting form must be of rank two as independent degrees of freedom are bi-dimensional (i.e.
 quantity plus conjugate).
@@
 \begin{equation}\label{mdof_form_antisymm}
-	\omega(\vec{v}, \vec{w}) = - \omega(\vec{w}, \vec{v}).
+	\omega(\vec{v}, \vec{w}) = - \omega(\vec{w}, \vec{v})
 \end{equation}
@@
 Suppose you take a surface and translate it along another independent variable.
-The count of states changed along the translation cannot depend on the independent variable.
-Therefore if we construct any parallelepiped at equal time, two opposite sides will contain the same number of states with opposite orientation, and therefore the integral of $\omega$ over the whole surface would be zero.
+The count of states cannot depend on the independent variable.
+Therefore if we construct any parallelepiped at equal time, two opposite sides will contain the same number of states with opposite orientation, and therefore the integral of $\omega$ over the whole surface will be zero.
 
@@
 \end{align}
-Mathematically we say that the displacement vector kills the form, it makes it zero no matter what vector we put on the other side.
+Mathematically we say that the displacement vector kills the form; it makes it zero no matter what vector we put on the other side.
 
@@
 \end{equation}
-If $\gamma$ is a solution of the equations of motions, then $\frac{d\gamma}{dt} = \vec{S}$.
-By \ref{mdof_displacement_kills}, $\vec{S}$ is the only vector field that kill the form $\omega$ therefore the solutions are such that
+If $\gamma$ is a solution of the equations of motion, then $d\gamma = \vec{S} dt$.
+The infinitesimal surface $d\Sigma$ will be a parallelogram formed by $d\gamma$ and the direction of the variation $d\lambda$.
+By \ref{mdof_displacement_kills}, $\vec{S}$ is the only vector field that kills the form $\omega$ therefore the solutions are such that
 \begin{equation}\label{mdof_stationary_action}
@@
 	\begin{aligned}
-		\delta \int_{\gamma} \theta(d\gamma) &= \delta \int^{t_2}_{t_1} p_i \frac{dq_i}{dt} - H dt \\
+		\delta \int_{\gamma} \theta(d\gamma) &= \delta \int^{t_2}_{t_1}\left( p_i \frac{dq_i}{dt} - H \right) dt \\
 		 &= \delta \int^{t_2}_{t_1}L(q^i, \dot{q}^i, t) dt = 0.
@@
 
-Note that the true physical content is in the displacement field $\vec{S}$ and the counting form $\omega$ whose properties descend from the assumption.
-Both the Lagrangian and the potential $\theta$ are not uniquely defined and are subject to a convenient, though still arbitrary, choice of gauge.
-The Lagrangian and the action, then, do not directly encode physical property of the system.
-Saying that ``nature chooses the path that minimizes the action'', therefore, is not really that insightful because such a quantity is not uniquely defined.
-On the other hand, the three equivalent assumptions give us a much more clear picture.
+Note that the true physical content is in the displacement field $\vec{S}$ and the counting form $\omega$ whose properties descend from the assumptions.
+The action, the Lagrangian and the potential $\theta$ are not uniquely defined and are subject to a convenient, though still arbitrary, choice of gauge.
+The Lagrangian and the action, then, do not directly encode physical properties of the system.
+Saying that ``nature chooses the path that minimizes the action'', therefore, does not provide clear insight.
+On the other hand, the three assumptions give us a clear physical picture.
 
@@
 Also note that the cases where the Lagrangian does not admit a Hamiltonian are exactly the cases where the action principle fails to yield a unique solution.
-In this sense, the action principle is better understood as a feature of Hamiltonian mechanics, instead of Lagrangian mechanics.
-
-As a final comment, note that privileging the Hamiltonian picture is more in line with quantum mechanics; the idea of $\omega$ as a tool to count states is in line with statistical mechanics; the expression \ref{mdof_potential_expression} is remarkably similar to that of the relativistic four-momentum.
+In this sense, the action principle is better understood as a feature of Hamiltonian mechanics, instead of Lagrangian mechanics.(TODO: check Arnold's book for reference)
+
+As a final comment, note that privileging the Hamiltonian picture is more in line with quantum mechanics; the idea of $\omega$ as a tool to count states is in line with statistical mechanics; expression \ref{mdof_potential_expression} is remarkably similar to that of the relativistic four-momentum.
 This is one of the key insights we get from our Reverse Physics approach: there is a unity among the different physical theories that begs to be brought to light.
@@
 
-We divide the appendix into three section: the first will address the single d.o.f.
+We divide the appendix into three sections: the first will address the single d.o.f.
 case, the second will give a geometric interpretation of the tools of differential geometry that is suitable for physicists, and lastly we address the multiple d.o.f.
@@
 For (i), determinism and reversibility means that giving the state at some point in time means constraining the state at all times, that the system can evolve in only one way.
-In math terms, if $X$ is the manifold of all possible states at times, given $P = \{q_0, p_0, t_0\}$ there will be only one possible evolution $\gamma : \mathbb{R} \to X$ such that $\gamma(t_0) = P$.
+In math terms, if $X$ is the manifold of all possible states at all times, given $P = \{q_0, p_0, t_0\}$ there will be only one possible evolution $\gamma : \mathbb{R} \to X$ such that $\gamma(t_0) = P$.
 
@@
 Equivalently, if the system has a given probability of starting in a given initial state, it must have the same probability of ending in the corresponding final state.
-In mathematical terms, if $\rho : X \to \mathbb{R}$ is a statistical distribution (or a probability density) defined on our extended space time and $\gamma$ is a possible evolution, we must have $\rho(\gamma(t_0)) = \rho(\gamma(t_1))$.
+In mathematical terms, if $\rho : X \to \mathbb{R}$ is a statistical distribution (or a probability density) defined on our extended phase space and $\gamma$ is a possible evolution, we must have $\rho(\gamma(t_0)) = \rho(\gamma(t_1))$.
 
 This requirement tells us two things.
-First, is that $X$ must be a differentiable manifold.
-To define densities over a set of variables and map them back and forward in time Jacobians must be defined, which means the derivative across state variables must be defined as well, in particular with respect with time, which plays the double role of a coordinate of $X$ and the evolution parameter.
-This means that the evolutions must be differentiable curves and therefore $\vec{S}$ must be defined which leads to
+First, that $X$ must be a differentiable manifold.
+To define densities over a set of variables and map them backward and forward in time Jacobians must be defined, which means the derivative across state variables must be defined as well, in particular with respect to time, which plays the double role of a coordinate of $X$ and the evolution parameter.
+This means that the evolutions must be differentiable curves and therefore the displacement $\vec{S}$ must be defined, which leads to
 \begin{equation}
@@
 Note that the vector potential is not uniquely identified, since $\nabla \times(\vec{\theta} + \nabla f) = \nabla \times \vec{\theta}$ and therefore the displacement field $\vec{S}$ remains unchanged.
-This is what is commonly referred to as gauge freedom in physics, and it is analogous to what happens for a magnetic field or for an incompressible fluid.
+This is commonly referred to as gauge freedom in physics, and it is analogous to what happens for a magnetic field or for an incompressible fluid.
 
@@
 
-We now use the constraint \ref{sdof_time_constraint} on the time component. We have
+We now use constraint \ref{sdof_time_constraint} on the time component. We have
 \begin{align}
@@
 Again note the double role of time: $\frac{dp}{dt}$ is not the same as $\frac{\partial p}{\partial t}$.\footnote{We could be more precise and use $\tau$ for the evolution parameter, leaving $t$ just for the coordinate.
-The treatment would automatically acquire a relativistic flavor and would enlighten us to other interesting results, though it would distract from the main purpose of this paper, the action principle.
+The treatment would automatically acquire a relativistic flavor and would enlighten us to other interesting results, though it would distract from the main focus of this paper, the action principle.
 } In the first case we are taking a total derivative along the evolution, and therefore the momentum can change.
@@
 
-Note that if we started from Hamilton's equation, the system would be deterministic and reversible: position and momentum at a fixed time would indeed allow us to pick a single differentiable evolution and, by Liouville's theorem, densities would be constant along such evolution.
-Therefore we have an equivalence between assumption $\ref{assum_detrev}$ and Hamilton's equations $\ref{sdof_Ham_eq}$ in the single degree of freedom case.
+Note that if we started from Hamilton's equations, the system would be deterministic and reversible: position and momentum at a fixed time would indeed allow us to pick a single differentiable evolution and, by Liouville's theorem, densities would be constant along such evolution.
+Therefore we have an equivalence between assumption \ref{assum_detrev} and Hamilton's equations \ref{sdof_Ham_eq} in the single degree of freedom case.
 
@@
 We take a path $\gamma$ and a variation $\gamma'$.
-Since they share the endpoints, taken together they form a closed line where $\gamma'$ is taken with the opposite orientation.
+Since they share the endpoints, taken together they form a closed curve where $\gamma'$ is taken with the opposite orientation.
 Because $\vec{S}$ is divergence-free, the flow of $\vec{S}$ through a surface depends only on its contour.
-Moreover, by Stoke's theorem, the flow through the surface will be equal to the integral of the potential over the contour.
+Moreover, by Stokes' theorem, the flow through the surface will be equal to the integral of the potential over the contour.
 Putting it all together we have
@@
 
-If $\gamma$ is an actual solution of the equation of motion, $\gamma$ is a field line of $\vec{S}$: $\gamma$ is tangent to $\vec{S}$ at all points.
+If $\gamma$ is a solution of the equations of motion, $\gamma$ is a field line of $\vec{S}$: $\gamma$ is tangent to $\vec{S}$ at all points.
 We can write $d\vec{\gamma} = \vec{S} dt$.
@@
 We can reason the other way: we look for a line $\gamma$ such that the triple product is zero regardless of what $d\vec{\lambda}$ is.
-The only way for that to happen is if $\gamma$ is tangent to $\vec{S}$ at each point, which means $\gamma$ is a solution of the equation of motion.
-This means that the paths that make the line integral of the vector potential stationary are those and only those that solve the equations of motions.
-This recovers the principle of the stationary action in Hamiltonian form
+The only way for that to happen is if $\gamma$ is tangent to $\vec{S}$ at each point, which means $\gamma$ is a solution of the equations of motion.
+This means that the paths that make the line integral of the vector potential stationary are those and only those that solve the equations of motion.
+This recovers the principle of stationary action in Hamiltonian form
 \begin{equation}
@@
 
-We can use this map to express the Hamiltonian form of principle of stationary action only in terms of position and velocity we have
+We can use this map to express the Hamiltonian form of the principle of stationary action only in terms of position and velocity. We have
 \begin{equation}
@@
 		&= \delta \int^{t_2}_{t_1} \left(p \frac{dq}{dt} - H\right) dt \\
-		&= \delta \int^{t_2}_{t_1} L(q, \dot{q}, t) \, dt
+		&= \delta \int^{t_2}_{t_1} L(q, \dot{q}, t) \, dt = 0
 	\end{aligned}
@@
 The system satisfies \ref{assum_kineq}.
-Moreover, we can find a unique Hamiltonian such that the solution of the equations of motion are exactly those paths that satisfy the principle of stationary actions.
+Moreover, we can find a unique Hamiltonian such that the solutions of the equations of motion are exactly those paths that satisfy the principle of stationary action.
 Therefore the system satisfies \ref{assum_detrev} as well.
 
-Before proceed to the generalizations, a few observations are in order.
+Before proceeding to the generalization, a few observations are in order.
 \begin{itemize}
-	\item We stress that the vector potential $\vec{\theta}$ in not uniquely defined by the physics, and therefore neither are the action and the Lagrangian.
-Giving a precise physical characterizations of those two objects, then, is hard.
-In our view, it is a hopeless task.
+	\item We stress that the vector potential $\vec{\theta}$ is not uniquely defined, and therefore neither are the action nor the Lagrangian.
+Giving precise physical characterizations of those two objects, then, is in our view a hopeless task.
 	\item We found that the Hamiltonian $H$ is the time component of the vector potential $\vec{\theta}$, not a scalar. This relativistic aspect emerges naturally out of the assumptions.
@@
 These tools are not typically taught in a standard physics curriculum.
-If they are, their presentation is dense and abstract even to some mathematicians: at a topology conference, a math graduate student confessed he sticked with topology because differential geometry was too abstract for him, comment that was taken with sympathy by others.
-
+If they are, their presentation is dense and abstract.
 Apart from the abstract definitions, the notation gets in the way of the physics.
@@
 
-What we will give here, then, is not a fully rigorous introduction to the subject, but rather a geometrical understanding of what these tool describe and a physical motivation as to why they are useful.
-We hope this will make it easier to follow the next section and to build a geometrical and physical intuition of the principle of stationary actions, which is the whole point of this article.
+What we will give here, then, is not a fully rigorous introduction to the subject, but rather a geometrical understanding of what these tools describe and a physical motivation as to why they are useful.
+We hope this will make it easier to follow the next section and to build a geometrical and physical intuition of the principle of stationary action, which is the focus of this article.
 
@@
 \end{equation*}
-The magnetic flux $\Phi$ through a surface $\Sigma$ can be understood as the sum of the contributions of the magnetic field $\vec{B}$ over each infinitesimal $d\Sigma$:
+The magnetic flux $\Phi$ through a surface $\Sigma$ can be understood as the sum of the contributions of the magnetic field $\vec{B}$ over each infinitesimal $\vec{d\Sigma}$:
 \begin{equation*}
@@
 \end{equation*}
-The work $W$ over a path $\lambda$ can be understood as the sum of the contributions of a force field $\vec{f}$ over each infinitesimal line $d\gamma$:
+The work $W$ over a path $\lambda$ can be understood as the sum of the contributions of a force field $\vec{f}$ over each infinitesimal line $\vec{d\gamma}$:
 \begin{equation*}
@@
 Though the dimensionality changes, a pattern emerges: the functionals $W$, $\Phi$ and $m$ all take a region of space while $\vec{f}$, $\vec{B}$ and $\rho$ act on infinitesimal regions.
-However, the vector representation gives us three different enough objects (i.e.
-vector, pseudo-vector and pseudo-scalar) and operations, and therefore fails fail to provide a nice generalization.
+However, the vector representation is in terms of different objects (i.e.
+vector, pseudo-vector and pseudo-scalar) and operations.
+Thus it fails to provide a nice generalization.
 
@@
 
-An infinitesimal region can also be understood as a parallelopiped, which is fully identified by its sides.
-Therefore a differential form can be understood as acting on a set of vectors, the sides of the parallelopiped, that matches the dimensionality of the form.
+An infinitesimal region can also be understood as a parallelepiped, which is fully identified by its sides.
+Therefore a differential form can be understood as acting on a set of vectors, the sides of the parallelepiped, that matches the dimensionality of the form.
 A one-form will take one vector, a two-form two vectors and so on.
-All forms must be anti-symmetric because switching the order of the sides does not change the parallelopiped, just its orientation.
+All forms must be anti-symmetric because switching the order of the sides does not change the parallelepiped, just its orientation.
 
 This picture is much easier to generalize to any dimensionality.
-If we note $S^k$ the space of all the $k$ dimensional subregions, a $k$-functional $F : S^k \to \mathbb{R}$ is a linear functional that takes a $k$-surface and returns a number.
+If we note $S^k$ the space of all the $k$ dimensional subregions, a $k$-functional $F : S^k \to \mathbb{R}$ is a linear functional that takes a $k$-surface $\sigma$ and returns a number.
 This can be expressed as
@@
 \end{equation*}
-the integral of a $k$-form $\omega : V^k \to \mathbb{R}$ that takes an infinitesimal $k$-surface (or equivalently a set of $k$ vectors) and returns a number.
+the integral of a $k$-form $\omega : V^k \to \mathbb{R}$ that takes an infinitesimal $k$-surface $d\sigma$ (or equivalently a set of $k$ vectors) and returns a number.
 
@@
 \end{equation*}
-These tensor are all anti-symmetric as the forms are anti-symmetric.
+These tensors are all anti-symmetric as the forms are anti-symmetric.
 
@@
 This is equated to the integral over the boundary of the field itself.
-But again, the vector calculus notation prevents us to generalize as the differential operator is somewhat different in each of the three cases.
+But again, the vector calculus notation prevents us from generalizing as the differential operator is somewhat different in each of the three cases.
 
@@
 However, its boundary $\partial \Sigma$ will be a $k$-dimensional surface, which is a valid argument for $F$.
-Therefore from any $F$ we can create a $k+1$-functional, which we can call exterior functional of $F$, noted $\partial F$, such that $\partial F(\Sigma) = F(\partial \Sigma)$.
+Therefore from any $F$ we can create a $k+1$-functional, which we can call the exterior functional of $F$, noted $\partial F$, such that $\partial F(\Sigma) = F(\partial \Sigma)$.
 If we look at all three cases above, that is exactly what is happening: the gradient, the curl and the divergence lead to a functional one dimension higher by applying the original operation on the boundary.
@@
 \end{equation*}
-This is the generalized Stokes' theorem and $\partial \omega$ is the exterior derivative. We use $\partial$ as we leave $d$ reserved for the infinitesimal regions.
-
-With a few steps, one can see that the divergence, curl and gradient correspond to the exterior derivative in three dimensions for zero-form, one-form and two-form respectively, and that the three theorems of vector calculus are particular instances of this generalized one.
+This is the generalized Stokes' theorem and $\partial \omega$ is the exterior derivative of $\omega$. We use $\partial$ as we leave $d$ reserved for the infinitesimal regions.
+
+With a few steps, one can see that the divergence, curl and gradient correspond to the exterior derivative in three dimensions for a zero-form, one-form and two-form respectively, and that the three theorems of vector calculus are particular instances of this generalized one.
 
@@
 	\begin{aligned}
-		\nabla \cdot \nabla \times \vec \vec{f} &= 0 \\
+		\nabla \cdot \nabla \times \vec{f} &= 0 \\
 		\nabla \times \nabla \varphi &= 0
@@
 We have  $\partial \partial F(\Sigma) = F(\partial \partial \Sigma)$ and the boundary of a boundary $\partial \partial \Sigma = \emptyset$ is always the empty set.
-So, no matter what $F$ or $\Sigma$ are, we are evaluating a functional on an empty set, which is zero.
-
-In the same way that a divergence-free field can be written as the curl of a vector potential, or a curl-free field can be written as the gradient of a scalar potential, a form whose exterior derivative $\partial \omega = 0$ is zero (a closed form), can be written as the exterior derivative $\omega = \partial \theta$ of a lower dimensional form, at least is some neighborhood.
+So, no matter what $F$ or $\Sigma$ is, we are evaluating a functional on an empty set, which is zero.
+
+In the same way that a divergence-free field can be written as the curl of a vector potential, or a curl-free field can be written as the gradient of a scalar potential, a form whose exterior derivative is zero $\partial \omega = 0$ (a closed form), can be written as the exterior derivative $\omega = \partial \theta$ of a lower dimensional form, at least in some neighborhood.
 In terms of functionals, this corresponds to the case where $F(\sigma)=0$ for all contractible surfaces, those surfaces that can be morphed to a point with a continuous transformation.
-If $F(\sigma)=0$ over all surfaces, instead, then the correspond is exact.
+If $F(\sigma)=0$ over all surfaces, instead, then the corresponding form is exact.
 
@@
 
-The case of multiple degrees of freedom proceeds in the same way as the single degree of freedom. We will have to:
+(TODO: Describe the space) The case of multiple degrees of freedom proceeds in the same way as the single degree of freedom. We will have to:
 \begin{enumerate}[label=(\roman*)]
@@
 We also need to reformulate state/density conservation in time, which we cannot do simply with a surface integral of $\vec{S}$.
-To this end, we need a functional $\mathcal{F}$ that takes an hyper-surfaces $\Sigma$ in the extended phase space and returns the count of states.
+To this end, we need a functional $\mathcal{F}$ that takes a hyper-surface $\Sigma$ in the extended phase space and returns the count of states.
 We have
@@
 
-As in the previous case, assumption \ref{assum_detrev} tells us that the integral over a closed hyper-surface is zero, since as many states flow in and out. We have
+As in the previous case, assumption \ref{assum_detrev} tells us that the integral over a closed hyper-surface is zero, since as many states flow in as flow out. We have
 \begin{equation}
@@
 
-Given that $\Omega$ quantifies the flow over $\vec{S}$, the flow for any surface that is tangent to $\vec{S}$ must vanish. If $d\gamma = \vec{S} \varphi dt$ for some scalar $\varphi$, we must have
+Given that $\Omega$ quantifies the flow of $\vec{S}$ through $\Sigma$, the flow for any surface that is tangent to $\vec{S}$ must vanish.
+If $d\gamma = \vec{S} \varphi dt$ for some scalar $\varphi$, we must have
 \begin{equation}
@@
 \end{equation}
-The notation is a bit improper as $\vec{S}$ is not technically an infinitesimal displacement, but a rate of displacement given a parameter (i.e.
-a velocity).
-However, the slight abuse of notation makes the concept more clear: the direction of motion does not contribute to the flow.
+The notation is a bit improper as $\vec{S}$ is not technically an infinitesimal displacement, but a rate of displacement given a parameter.
+However, the slight abuse of notation makes the concept more clear: the direction of motion does not contribute to the flow through the surface $\Sigma$.
 
@@
 We now turn to assumption \ref{assum_indep}.
-This requires the count of states to factorize: if we have a parallelopiped in phase space, the count of states (i.e.
-its volume) is the product of the count of the configurations over each initial conditions (i.e.
+This requires the count of states to factorize: if we have a parallelepiped in phase space, the count of states (i.e.
+its volume) is the product of the count of the configurations over each independent d.o.f.
+(i.e.
 the area of the sides).
-Equivalently, \ref{assum_indep} must allow the case of a distribution over states with no correlations between the independent d.o.f., in which cases the density must factorize.
+Equivalently, \ref{assum_indep} must allow the case of a distribution over states with no correlations between the independent d.o.f., in which case the density must factorize.
 Since we must be able to quantify the configuration for each d.o.f., and since each d.o.f.
@@
 If we imagine translating in time, again the number of configurations cannot change because of \ref{assum_detrev}.
-Therefore, if we imagine a parallelopiped, opposite sides must identify the same number of configurations.
+Therefore, if we imagine a parallelepiped, opposite sides must identify the same number of configurations.
 The integral will take the sides with an opposite orientation, and therefore there will be zero net contribution.
@@
 \begin{equation}
-	\partial \Omega = \partial \omega \wedge \omega^{N-1} = 0.
+	\partial \Omega = N\partial \omega \wedge \omega^{N-1} = 0.
 \end{equation}
@@
 \end{equation}
-If we start with a closed form $\omega$ that has a single degenerate direction, by Darboux theorem we can find coordinates for the non-degenerate part $\omega$ that satisfy the above conditions.
+If we start with a closed form $\omega$ that has a single degenerate direction, by Darboux's theorem we can find coordinates for the non-degenerate part of $\omega$ that satisfy the above conditions.
 This recovers \ref{assum_indep}.
 
-Note that we have broken down exactly how each assumptions contributes to the mathematical setting.
-If \ref{assum_indep} fails, then $\Omega$ cannot be understood as $\omega^N$, which means we have much fewer condition on the motion: we only know that the total flow is conserved, but we cannot break it down along each d.o.f..
+Note that we have broken down exactly how each assumption contributes to the mathematical setting.
+If \ref{assum_indep} fails, then $\Omega$ cannot be understood as $\omega^N$, which means we have much fewer conditions on the motion: we only know that the total flow is conserved, but we cannot break it down along each d.o.f..
 On the other hand, if \ref{assum_detrev} fails we may not have a single displacement field $\vec{S}$ and/or the form that counts states would not be closed.
@@
 
-The insight is that the independence of the degree of freedom leads to the independent expression of the components of the covector potential.
+The insight is that the independence of the degrees of freedom leads to the independent expression of the components of the covector potential.
 We stress again that this is a choice, and that the specific expression of the components is not physically constrained.
@@
 
-For (iii) we calculate $\omega(S, \cdot ) $. We have
-\begin{equation}
-\begin{aligned}
-	\omega(S, \cdot )  &= S^a \omega_{ab} e^b = 0 \\
+For (iii) we calculate $\omega(\vec{S}, \cdot ) $. We have
+\begin{equation}
+\begin{aligned}
+	\omega(\vec{S}, \cdot )  &= S^a \omega_{ab} e^b = 0 \\
 	&= (S^{q^i}\omega_{q^ib} + S^{p_i}\omega_{p_ib} + S^{t}\omega_{tb}) e^b \\
@@
 For those familiar with general relativity, recall that the metric tensor $g_{\alpha\beta}$ defines the geometry by defining lengths and angles.
-Here we something similar in the sense that the two-form $\omega_{ab}$ is a rank two tensor that defines the geometry of the space.
-The difference is that $\omega_{ab}$ does not define lengths and angles, but only areas of two-dimensional surfaces, which corresponds to the number of configuration on each independent degree of freedom.
+Here we find something similar in the sense that the two-form $\omega_{ab}$ is a rank two tensor that defines the geometry of the space.
+The difference is that $\omega_{ab}$ does not define lengths and angles, but only areas of two-dimensional surfaces, which corresponds to the number of configurations on each independent degree of freedom.
 Mathematically, this difference is between the symmetry of $g_{\alpha\beta} = g_{\beta\alpha}$ and the anti-symmetry of $\omega_{ab} = -\omega_{ba}$.
@@
 This is the same relationship that we would have, for example, for the magnetic field psuedo-vector and the spatial components of the electromagnetic field in relativity.
-If we express $\omega$ in terms of $\theta$, we have
+If we express $\omega$ in terms of $\vec{\theta}$, we have
 \begin{equation}
@@
 For (iv), we simply need to apply Stokes' theorem.
-The situation is identical to the one degree of freedom, except that the flow now is not given by the surface integral of $\vec{S}$, but of the form $\omega$.
-The path $\gamma$ and its variation $\gamma'$ will form a closed line where $\gamma'$ is taken with the opposite orientation.
+The situation is identical to the single degree of freedom, except that the flow now is not given by the surface integral of $\vec{S}$, but of the form $\omega$.
+The path $\gamma$ and its variation $\gamma'$ will form a closed curve where $\gamma'$ is taken with the opposite orientation.
 Because $\omega$ is closed, its surface integral depends only on its contour.
-Moreover, by Stoke's theorem, the flow through the surface will be equal to the integral of the potential over the contour.
-Putting it all together we recover \ref{mdof_action}
-\begin{equation}
+Moreover, by Stokes' theorem, the flow through the surface will be equal to the integral of the potential over the contour.
+Putting it all together we recover
+\begin{equation}
+\tag*{(\ref{mdof_action})}
 \begin{aligned}
@@
 	\int_{\Sigma} \partial \theta (d\Sigma) \\
-	&= \oint_{\partial \Sigma} \theta(d\partial\Sigma) \\
+	&= \oint_{\partial \Sigma = \gamma \cup \gamma'} \theta(d\gamma) \\
 	&= \int_{\gamma} \theta(d\gamma) - \int_{\gamma'} \theta (d\gamma') \\
@@
 
-For (v), if the path $\gamma$ is a solution to the equation of motion, it is tangent at every point to $\vec{S}$.
+For (v), if the path $\gamma$ is a solution to the equations of motion, it is tangent to $\vec{S}$ at every point.
 Therefore we can write $d\gamma= \vec{S} dt$.
-Each infinitesimal surface element $d\Sigma$ between the line and the variation will have $d\gamma$ as one side.
-Let us casll $d\lambda$ the other side, which is the displacement that takes the path to the variation.
+Each infinitesimal surface element $d\Sigma$ will be an infinitesimal parallelogram with $d\gamma$ as one side, and the other side $d\lambda$ connects the path to the variation.
 We have:
 \begin{equation}
-\begin{aligned}
-	-\int_\Sigma \omega(d\Sigma) &= -\int_\Sigma \omega(d\gamma, d\lambda) \\
+\tag*{(\ref{mdof_stationary_action})}
+\begin{aligned}
+	\delta \int_{\gamma} \theta (d\gamma) &= -\int_\Sigma \omega(d\Sigma) \\
+	&= -\int_\Sigma \omega(d\gamma, d\lambda) \\
 	&= -\int_\Sigma \omega(\vec{S}, d\lambda) dt = 0.
@@
 \end{equation}
-The converse is also true: suppose the flow through $\gamma$ and $\gamma'$ is zero no matter the choice of gamma, this means that at every point $\omega(d\gamma, \cdot) dt = 0$, which means $d\gamma$ kills $\omega$.
+The converse is also true: suppose that, for a fixed $\gamma$, the flow through $\Sigma$ is zero no matter the choice of $\gamma'$.
+Then at every point $\omega(d\gamma, \cdot) dt = 0$, which means $d\gamma$ kills $\omega$.
 But since $\vec{S}$ is the only direction that kills $\omega$ at every point, $d\gamma$ must be everywhere tangent to $\vec{S}$ and therefore $\gamma$ is a solution of the equations of motion.
@@
 \begin{equation}
+\tag*{(\ref{mdof_Lagrangian})}
 	\begin{aligned}
 		\delta \int_\gamma \theta (d\gamma) &= \delta \int^{t_2}_{t_1} \theta \left(\frac{d\gamma}{dt} dt\right) \\
-		&= \delta \int^{t_2}_{t_1} \left(\theta_{q^i} \frac{dq^i}{dt} + \theta_{p_i} \frac{dp^i}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
+		&= \delta \int^{t_2}_{t_1} \left(\theta_{q^i} \frac{dq^i}{dt} + \theta_{p_i} \frac{dp_i}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
 		&= \delta\int^{t_2}_{t_1} \left(p_i \frac{dq^i}{dt} + 0 \frac{dp_i}{dt} - H \frac{dt}{dt}\right) dt \\
 		&= \delta \int^{t_2}_{t_1} \left(p_i \frac{dq^i}{dt} - H\right) dt \\
-		&= \delta \int^{t_2}_{t_1} L(q^i, \dot{q}^i, t) \, dt.
+		&= \delta \int^{t_2}_{t_1} L(q^i, \dot{q}^i, t) \, dt = 0.
 	\end{aligned}
@@
 \begin{equation}\label{mdof_potential_relativistic}
-	\theta = p_i e^{q^i} - \frac{H}{c} e^{ct}.
+	\theta = p_i e^{q^i} - \frac{H}{c} e^{ct},
 \end{equation}
 which bears striking resemblance to the relativistic four-momentum.
-While it goes outside the scope of this article, the idea is that, in this setting, we already have some pre-relativistic features, even though we have not introduced a metric tensor.
+While it is outside the scope of this article, the idea is that, in this setting, we already have some pre-relativistic features, even though we have not introduced a metric tensor.
+
+We hope that this result shows that it is possible to bring physical ideas and mathematical concepts much closer together in a way that is more clear and satisfying.
 

### 2022-09-28 Articles/2022-ActionPrinciple/ActionPrincipleArchive.tex
ActionPrinciple: uploaded to arXiv

--- 2022-08-07 TeachingMaterial/ActionPrinciple/ActionPrincipleArticle.tex
@@
 
-While the principle of stationary action is regarded by many as one of the most important tools in physics, its physical meaning is not completely clear.
+While the principle of stationary action is regarded by many as one of the most important tools in physics,\cite{hamilton1834general, heaviside1903, feynman1942principle, brenier1989least, rojo2018principle} its physical meaning is not completely clear.
 First of all, the typical characterization of the Lagrangian as the difference between kinetic and potential energy fails even for simple systems, like a charged particle under a magnetic field.
@@
 
-As part of our larger project Assumptions of Physics, we developed an approach, called Reverse Physics (TODO cite paper), which examines current theories to find a set of starting physical assumptions that are sufficient to rederive them.
-We have found that Lagrangian mechanics is equivalent to three assumptions: determinism/reversibility, independence of degrees of freedom and kinematics/dynamics equivalence (TODO cite long paper).
+As part of our larger project Assumptions of Physics, we developed an approach, called Reverse Physics \cite{aop-phys-ReversePhysics}, which examines current theories to find a set of starting physical assumptions that are sufficient to rederive them.
+We have found that Lagrangian mechanics is equivalent to three assumptions: determinism/reversibility, independence of degrees of freedom and kinematics/dynamics equivalence \cite{aop-phys-blueprint}.
 This physically motivated understanding of the classical theory can be used to characterize both the physics and the geometry underlying the principle of stationary action.
@@
 
-The mathematics needed to run the argument is not new\cite{souriau1997structure}.
+The mathematics needed to run the argument is well established\cite{souriau1970structure, abraham1978foundations,arnold1989mathematical, marsden1999introduction}.
 The two aspects that have been sorely lacking, and that we provide, are (1) a clear geometrical interpretation of the action principle and (2) a tight connection between the math and the physics it represents.
@@
 
-As we want to characterize the evolution of states over time, the appropriate setting is phase space extended with the time variable.(TODO find citations) That is, the space charted by position $q$, momentum $p$ and time $t$ as can be seen in Fig.
+As we want to characterize the evolution of states over time, the appropriate setting is phase space extended with the time variable\cite{lanczos1949variational,synge1960encyclopedia}.
+That is, the space charted by position $q$, momentum $p$ and time $t$ as can be seen in Fig.
 \ref{extended_phase_space}.
@@
 \end{equation}
-which we call the state counting form.\footnote{Note that this approach is consistent with statistical mechanics. (TODO: add references)}
+which we call the state counting form.\footnote{Note that this approach is consistent with statistical mechanics\cite{peliti2011statistical}.}
 
@@
 Also note that the cases where the Lagrangian does not admit a Hamiltonian are exactly the cases where the action principle fails to yield a unique solution.
-In this sense, the action principle is better understood as a feature of Hamiltonian mechanics, instead of Lagrangian mechanics.(TODO: check Arnold's book for reference)
+In this sense, the action principle is better understood as a feature of Hamiltonian mechanics, instead of Lagrangian mechanics\cite{souriau1970structure, arnold1989mathematical}.
 
@@
 
-(TODO: Describe the space) The case of multiple degrees of freedom proceeds in the same way as the single degree of freedom. We will have to:
+The case of multiple degrees of freedom proceeds in the same way as the single degree of freedom.
+The main difference is that we will work with a system described by $N$ degrees of freedom, which is therefore described by the $2N+1$ variables
+\begin{equation}
+\tag*{(\ref{mdof_variables})}
+	\xi^a = [ q^i, p_i, t].
+\end{equation}
+We will have to:
 \begin{enumerate}[label=(\roman*)]

### 2022-10-13 Articles/2022-ActionPrinciple/ActionPrincipleArchive.tex
ActionPrinciple: adding note on nonholonomic constraints
ActionPrinciple: geometrical -> geometric

@@
 
-\title{Geometrical and physical interpretation of the action principle}
+\title{Geometric and physical interpretation of the action principle}
 \author{Gabriele Carcassi, Christine A. Aidala}
@@
 \begin{abstract}
-We give a geometrical interpretation for the principle of stationary action in classical Lagrangian particle mechanics.
+We give a geometric interpretation for the principle of stationary action in classical Lagrangian particle mechanics.
 In a nutshell, the difference of the action along a path and its variation effectively ``counts'' the possible evolutions that ``go through'' the area enclosed.
@@
 The mathematics needed to run the argument is well established\cite{souriau1970structure, abraham1978foundations,arnold1989mathematical, marsden1999introduction}.
-The two aspects that have been sorely lacking, and that we provide, are (1) a clear geometrical interpretation of the action principle and (2) a tight connection between the math and the physics it represents.
+The two aspects that have been sorely lacking, and that we provide, are (1) a clear geometric interpretation of the action principle and (2) a tight connection between the math and the physics it represents.
 To make the result accessible to the widest audience, we first cover the case of a single degree of freedom using standard vector calculus.
 We then proceed to the case of multiple independent degrees of freedom using tools from differential geometry.
-The main article will presents all the key point needed to follow the argument and its physical and geometrical meaning, leaving the mathematical details and calculations to the appendix.
+The main article will presents all the key point needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the appendix.
 
@@
 
-We have thus demystified the action principle, and turned it into a geometrical property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field in phase space.
+We have thus demystified the action principle, and turned it into a geometric property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field in phase space.
 We also have a clear physical meaning: the principle of stationary action is equivalent to assuming determinism/reversibility \ref{assum_detrev} and kinematic equivalence \ref{assum_kineq}.
@@
 The exterior derivative takes the place of the curl and it is $\omega$, not $\vec{S}$, that can be expressed in terms of the potential $\theta$.
-Yet, there is an important geometrical relationship we still need to take into account: the direction of the displacement does not contribute new states.
+Yet, there is an important geometric relationship we still need to take into account: the direction of the displacement does not contribute new states.
 Physically, this is a consequence of \ref{assum_detrev}: states cannot be created or disappear over time.
@@
 We have found that the principle of stationary action is equivalent to assumptions \ref{assum_detrev}, \ref{assum_indep} and \ref{assum_kineq} and that the variation of the action can be understood as the flow of evolutions through the surface delimited by the path and its variation.
-This allows us to fully understand both the physical and geometrical significance of the principle in a much more precise way.
+This allows us to fully understand both the physical and geometric significance of the principle in a much more precise way.
 
@@
 This is one of the key insights we get from our Reverse Physics approach: there is a unity among the different physical theories that begs to be brought to light.
-We are convinced that a version of this geometrical understanding must exist in the quantum world.
+We are convinced that a version of this geometric understanding must exist in the quantum world.
 
@@
 no states are created or destroyed).
-That is the whole physical and geometrical content of those equations.
+That is the whole physical and geometric content of those equations.
 
@@
 
-Now that we have developed a clear geometrical and physical understanding of Hamilton's equations, we can concentrate on the action principle.
+Now that we have developed a clear geometric and physical understanding of Hamilton's equations, we can concentrate on the action principle.
 For (iv) we can apply Stokes' theorem.
@@
 Similarly, we would find that the streamlines of an incompressible fluids are the lines and only the lines for which the integral of the vector potential of the flow is stationary.
-While these properties may be of limited practical use, the geometrical insight is the same.
+While these properties may be of limited practical use, the geometric insight is the same.
 
@@
 
-What we will give here, then, is not a fully rigorous introduction to the subject, but rather a geometrical understanding of what these tools describe and a physical motivation as to why they are useful.
-We hope this will make it easier to follow the next section and to build a geometrical and physical intuition of the principle of stationary action, which is the focus of this article.
+What we will give here, then, is not a fully rigorous introduction to the subject, but rather a geometric understanding of what these tools describe and a physical motivation as to why they are useful.
+We hope this will make it easier to follow the next section and to build a geometric and physical intuition of the principle of stationary action, which is the focus of this article.
 
@@
 
-%In our opinion, , and In order to fully understand the geometrical and physical meaning of the general case, it helps to have a crisp intuition for the tools of differential topology.
+%In our opinion, , and In order to fully understand the geometric and physical meaning of the general case, it helps to have a crisp intuition for the tools of differential topology.
 We will therefore go through the basic motivation of differential topology from a physicist perspective, so that we can see what differential forms and exterior derivatives are capturing at a a physical level.
@@
 On the other hand, if \ref{assum_detrev} fails we may not have a single displacement field $\vec{S}$ and/or the form that counts states would not be closed.
+Nonholonomic constraints will, in general, make one of the assumptions fail, for example by introducing a dependence between degrees of freedom.
 This type of exact mapping between assumptions and their mathematical expression is one of the key goals of Reverse Physics in particular and our overall project Assumptions of Physics more in general.

### 2023-01-21 Notes/2023-QuantumGeneralizedCoordinates/QuantumGeneralizedCoordinates.tex
Notes on variable changes in QM

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{enumitem}

\def\>{\rangle}
\def\<{\langle}

\begin{document}

\title{Notes on QM in generalized coordinates}
\author{Gabriele Carcassi}

\date{\today}

\maketitle

\begin{abstract}
\end{abstract}

\section{Change of coordinates in manifolds}

\begin{equation}
	\begin{aligned}
		\hat{x}^i &= \hat{x}^i(x^j) \\
		d\hat{x}^i &= \partial_j \hat{x}^i dx^j \\
		d\hat{x}^n &= | \partial_j \hat{x}^i | dx^n = |J| dx^n \\
		|J| \hat{\rho}(\hat{x}^i) &= | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i)  = \rho(x^j) \\
		\int \rho(x^j) dx^n &= \int | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i) dx^n = \int \hat{\rho}(\hat{x}^i) d\hat{x}^n
	\end{aligned}
\end{equation}

Mathematically, $\rho$ can be understood as the (only) component of a volume form. In that case, $\rho$ changes sign under change of orientation. Alternatively, $\rho$ can be understood as the (only) component of a density.\footnote{Effectively, the density is like a volume form that keeps track of orientation.} In this case, $\rho$ does not change sign under change of orientation and the transformation rule should use the absolute value of the Jacobian determinant.

\section{Change of coordinates for wave-function}
\begin{equation}
	\begin{aligned}
		\rho(x^j) &= \psi^\dagger(x^j) \psi(x^j)\\
		||J|| \hat{\rho}(\hat{x}^i) &= || \partial_j \hat{x}^i || \hat{\psi}^\dagger(\hat{x}^i) \hat{\psi}(\hat{x}^i) = \psi^\dagger(x^j) \psi(x^j) = \rho(x^j) \\
		\sqrt{|| \partial_j \hat{x}^i ||} \hat{\psi}(\hat{x}^i) &= \psi(x^j) \\
		\hat{\psi}(\hat{x}^i) &= \sqrt{|| \partial_i x^j ||} \psi(x^j) \\
	\end{aligned}
\end{equation}
Note that $\rho$ can't be negative. Mathematically,  $\rho$ should be the component of a density. Therefore $||J||$ represents the absolute value of the Jacobian determinant.

\section{Change of coordinates in symplectic manifolds}
\begin{equation}
	\begin{aligned}
		\hat{p}_i &= \partial_i x^j p_j \\
		d \hat{p}_i &= \partial_i x^j d p_j \\
		d\hat{p}^n &= | \partial_i x^i | dp^n = |J^{-1}| dx^n \\
	\end{aligned}
\end{equation}

\section{Properties of Fourier transform}

Definition of Fourier transform
\begin{equation}
	\begin{aligned}
		\phi(p_j) &= F(\psi(x^i)) = \frac{1}{\sqrt{2 \pi \hbar}} \int e^{\frac{p_j x^j}{\imath \hbar}} \psi(x^j) dx^n \\
	\end{aligned}
\end{equation}

Definition of convolution
\begin{equation}
	\begin{aligned}
		(g * h)(t) &= \int_{0}^{t} g(\tau) h(t-\tau) d\tau\\
	\end{aligned}
\end{equation}

Convolution theorems
\begin{equation}
	\begin{aligned}
		F((g * h)(t)) &= F(g(t)) F(h(t))\\
		F((g h)(t)) &= F(g(t)) * F(h(t))\\
	\end{aligned}
\end{equation}
TODO: find explicit confirmation for complex functions

\section{Change of Fourier transform}
\begin{equation}
	\begin{aligned}
		F(\hat{\psi}(\hat{x}^i)) &= F\left(\sqrt{|| \partial_i x^j ||} \psi(x^j)\right) = F\left(\sqrt{|| \partial_i x^j ||}\right) * F \left( \psi(x^j)\right)  \\
	\end{aligned}
\end{equation}
TODO: think of cross-checks

\end{document}

### 2023-01-21 Notes/2023-QuantumGeneralizedCoordinates/QuantumGeneralizedCoordinates.tex
More details on assumptions used

@@
 \section{Change of coordinates in manifolds}
-
+Transformation rules for differentials on a generic manifold. Only differentiable structure is required.
 \begin{equation}
@@
 		d\hat{x}^n &= | \partial_j \hat{x}^i | dx^n = |J| dx^n \\
+	\end{aligned}
+\end{equation}
+
+Transformation rules for a density $\rho(x^i)$. The requirement is that the integral is invariant.
+\begin{equation}
+	\begin{aligned}
+		\int \rho(x^j) dx^n &= \int | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i) dx^n = \int \hat{\rho}(\hat{x}^i) d\hat{x}^n \\
 		|J| \hat{\rho}(\hat{x}^i) &= | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i)  = \rho(x^j) \\
-		\int \rho(x^j) dx^n &= \int | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i) dx^n = \int \hat{\rho}(\hat{x}^i) d\hat{x}^n
 	\end{aligned}
@@
 \section{Change of coordinates for wave-function}
+
+Deducing transformation rules for wave-function based on the fact that the modulues of a wave-function is a density over the manifold.
 \begin{equation}
@@
 
-\section{Change of coordinates in symplectic manifolds}
+\section{Change of coordinates for momentum}
+
+These are the change of coordinates for momentum.
+They are derived by making an arbitrary change on the position coordinates and inducing a coordinate change on momentum such that the symplectic form $\omega = dx^i \wedge dp_i$ is invariant.
+Therefore this depends on the symplectic structure.
+This makes conjugate momentum vary as a covector.
+
 \begin{equation}
@@
 		d\hat{p}^n &= | \partial_i x^i | dp^n = |J^{-1}| dx^n \\
+	\end{aligned}
+\end{equation}
+
+Since in quantum mechanics we do not have a symplectic structure, we can find the same transformation rule based on the fact that $P_i = -\imath \hbar \partial_i$
+\begin{equation}
+	\begin{aligned}
+		\hat{P}_i &= -\imath \hbar \partial_i = -\imath \hbar \partial_i x^j \partial_j = \partial_i x^j P_j \\
 	\end{aligned}
@@
 \end{equation}
-TODO: think of cross-checks
+This expression depends only on the definition and properties of the Fourier transform and the transformation rules of densities of manifolds.
+
+TODO: Need a cross check with the transformation rule of momentum
 

### 2023-01-23 Notes/2023-QuantumGeneralizedCoordinates/QuantumGeneralizedCoordinates.tex
Adding ideas for cross check

@@
 
-TODO: Need a cross check with the transformation rule of momentum
+TODO: Need a cross check with the transformation rule of momentum. For example, show that the inner product in momentum basis is invariant:
+\begin{equation}
+	\begin{aligned}
+		\int \phi^\dagger(p_j) \phi(p_j) dp^n &= \int \hat{\phi}^\dagger(\hat{p}_i) \phi(\hat{p}_i) d\hat{p}^n  \\
+	\end{aligned}
+\end{equation}
+Another idea, show that applying the momentum operator in the new coordinates is equal to applying the the one in the old coordinates and then rescaling by the square root of the Jacobian.
 

### 2023-01-23 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
Revised draft

--- 2021-08-13 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
@@
 It fails simply because experiments do not agree with it.\footnote{CITE: papers that either make the claim, or that discuss how everybody makes that claim} This view is so established, so ingrained, that it is seldom discussed explicitly, and we believe it to be part of the reason why quantum mechanics still feels, after almost a century, so mysterious: how are we going to conceptually appreciate a theory if we do not even know what conceptual problems it is trying to address?
+[TODO: expand if needed]
 
-Our argument can be schematically summarized as follows:
+We present two arguments, which can be understood as the same argument in different form:
 \begin{enumerate}
-	\item To define a system, we need to define a clear boundary between the system and everything else
-	\item The interaction across the boundary participates in the definition of the system
-	\item Perfect characterization of a system requires perfect isolation
-	\item As we get to smaller and smaller parts, perfect isolation becomes untenable
+	\item The entropy of a single microstate over a continuum has minus infinite entropy, which makes classical mechanics incompatible with thermodynamics
+	\item Classical mechanics implicitly assumes perfect isolation at all scales, which can only be understood as a simplifying assumption
 \end{enumerate}
+The link between the two is that the zero entropy level, the lowest physically possible, corresponds (on the continuum) to a minimal uncertainty.
 In a nutshell, classical mechanics fails because the definition of a system is too simplistic.
@@
 
-We will analyze these ideas first conceptually and how they translate into inconsistencies between classical mechanics on one side and thermodynamics/statistical mechanics on the other.
-We will then show how quantum mechanics addresses some of these issues and how it leaves some open, to be solved in future theories.
+\section{Entropic minimum}
 
-\section{System/environment boundary}
+We will start our investigation with a seemingly innocent question.
+What is the entropy of a single classical microstate?
+That is, suppose we have a system composed of $N$ classical particles.
+A complete configuration, a microstate, is given by the position and momentum of all $N$ particles.
+Suppose we pick a macrostate that coincides with a single microstate, a pure microstate.
+What is the entropy associated to that pure microstate?
 
-When defining a physical system, we need to define, implicitly or explicitly, its boundary.
-Whether we are studying a solid object, an open container of gas, a chemical compound or a single particle, we need, at the very least, to be clear as to what is in the system and what is out.
-This much should be evident.
-What may be less evident is that the nature of the interaction at the boundary between the system and the environment is also important, not just in thermodynamics and statistical mechanics, but more broadly to all physical systems.
-Let us start with the following simple example.
-
-Suppose we want to study the motion of a cannon ball.
-As the cannon ball moves, air molecules will scatter off its surface.
-Yet, the effect produced will be negligible.
-We can therefore describe the state of the cannon ball with a single, infinitely precise, pair of position and momentum.
-Now suppose we want to study the motion of a speck of dust.
-This time the effect of the air molecules will not be negligible.
-We will therefore describe the state of the speck of dust with a probability distribution over position and momentum.
-Now suppose we want to study the motion of the cannon ball on the surface of the sun.
-Plasma will scatter off the ball catastrophically, destroying the ball.
-
-The example illustrates that the system, the interaction at the boundary and the state space for the system are not three independent choices.
-One may constrain the others.
-A cannonball at 5000 Kelvin does not exist, or at least not for long.
-Conceptually, we can think of a property as being intrinsic to the system only if the interaction with the environment leaves it, at least to a good approximation, undisturbed.
-If it is not left undisturbed, then it is no longer a property of the system itself: it is a shared property between the system and the environment because the nature of both comes into play in its definition.\footnote{CITE: maybe someone in complex systems, emergent properties, ...
-make similar claims}
-
-While we are used to think this way for thermodynamic quantities, like temperature or pressure, or statistical quantities, like average energy or mean free path, we typically think of mechanical properties, like position and velocity, differently.
-These, we think, are always intrinsic to the system.
-However, we have no basis for that distinction.
-What happens is that, in the vast number of cases, we can sufficiently isolate the system so that the effect of the environment on the motion of the body is negligible.
-We can take a speck of dust, put it in a vacuum, and the position is now much more stably definable.
-Therefore, when studying the speck in a fluid, the motion is modeled as fluctuations of an infinitely precise quantity because, at least conceptually, we could instantly remove the air and determine the position of the speck in isolation.
-But, realistically, we can only go so far.
-Perfect isolation is impossible.
-
-Apart from practical concerns, any wall we will use as shielding will necessarily have finite temperature and will emit thermal radiation; moreover, gravitational interactions cannot be shielded in the first place.\footnote{CITE: something that goes more into the details as to why gravitational interactions cannot be shielded} Even theoretically, complete isolation is not possible.
-We cannot, even conceptually, instantly remove all the background interactions.
-If that is the case, we have to concede that even properties like position and momentum are not intrinsic to the system: they are the characterization of the system together with the environment (e.g.
-the gravitational field).
-
-One may still argue that these are still circumstantial problems that derive from the incompleteness of our current understanding.
-We could always imagine that we will be able, in the future, to achieve perfect isolation.
-But let us explore that possibility in detail.
-Perfect isolation would mean that no signal, electromagnetic, gravitational or otherwise, would be able to interact with the object.
-Then, how are we supposed to measure anything about the system?
-It is experimentally inaccessible.
-Maybe we are already surrounded by perfectly isolated systems: we have no way of knowing.
-
-This issue uncovers another couple of hidden assumptions.
-First of all, the observer is always part of the environment as it is not part of the system.
-The second is that a measurement has to cross the boundary.
-The act of measurement may interfere with the previous nature of the system/environment interaction, which as we saw participates in defining the properties of the system itself.
-Therefore the act of measurement may change the nature of the very properties that we are going to measure.
-
-Note that these are definitional issues, not mechanistic issues.
-It is not about what the system is doing, what the rules are that the most fundamental objects follow.
-It is about how we define the system in the first place: how we carve up the world into parts that we can study independently.
-If we take a physical system simply as something we want to study under a set of conditions, and the state to be the experimentally accessible information that is well defined under such conditions, these issues will always crop up.
-They need to be addressed, even if just by simplifying assumptions, when the system is being defined.
-
-\section{Classical failure}
-
-We are now in a position to understand the failure of classical theories.
-As we saw before, we can assign intrinsic properties to systems that can be, at least in principle, isolated perfectly.
-For large scale objects, this is not a problem: small perturbations will have small effects which can, at least at a certain scale, be neglected.
-But this is an idealization.
-The infinite precision of a real number is valid only insofar that the idealization holds.
-
-Assuming perfect isolation by itself is not an issue.
-However, classical mechanics also assumes, implicitly, that this perfect isolation holds for objects of all sizes.
-With continuous materials, we can take smaller and smaller parts.
-Classical fields allow for wave packets of ever decreasing energy.
-The theory implicitly assumes all these objects can be prepared, manipulated and detected independently enough that their state can be perfectly defined.
-This is the contradiction inherent to classical mechanics: perfect isolation at all scales.
-
-The disconnect is that many take the classical state to be the ``real'' state of the system.
-But this is more of a philosophical construct.
-What physics should be concerned with is only the part of reality that is experimentally accessible under suitable conditions (e.g.
-repeatability, independence from other systems, ...).
-This does not mean that objects we cannot study experimentally do not exist or are not worth studying.
-It just means that our physical theories should not make any claims about them as they are not the subject of physics.
-
-Since this is a point of contention,\footnote{CITE: papers that show different view, papers that claim there are different views.} and we do not believe further discussion along these lines will change anyone's perspective, let us switch gears.
-We are going to argue that a perfect description of the type provided by classical mechanics is inconsistent with thermodynamics and statistical mechanics.
-
-We start by asking: what is the entropy of a single microstate?
-This is complicated by the fact that we have (at least) two definitions of entropies, and people argue about which is the correct one.\footnote{CITE: papers with different stances on entropy; papers that claim different stances} Therefore let us use both definitions and see what happens.
-
-Recall that a microstate for a classical system of $n$ particles is a point in a $\mathbb{R}^{6n}$ space, that represents all their three-dimensional position and momenta.
-The fundamental postulate of statistical mechanics states that the entropy is $S = k \log W$, where $W$ measures the  microstates.
-At first glance, since we have a single microstate, we are inclined to say $W = 1$.
-This would be true for a discrete state space.
-However, for a continuum state space this is the wrong answer.\footnote{This is an example where the difference between discrete and continuous spaces plays a fundamentally important role in physics.
-Unfortunately, too many physicists confuse discrete with countable.} We need to use the phase space volume.
-Since we have a single point, its volume is zero, therefore we have $S = k \log 0 = -\infty$: the entropy of a single microstate is minus infinity.
-
-Let us now confirm the result with the Gibbs/Shannon entropy $S = -\int \rho \log \rho \, dq^{3n} dp^{3n}$.
-A single microstate would be described by a delta distribution centered on that microstate: $\delta(q^i-q^i_0, p^i-p^i_0)$.
+Many would instinctively think that the answer is zero, but it not the case.
+Zero would be the correct answer if the space were discrete, but classical phase space is not discrete.
+To calculate the right answer, we can use the fundamental postulate of statistical mechanics\footnote{CITE} which states that the entropy is $k \log W$, where $W$, the count of microstates, is given by the volume of the region of phase space that represents the macrostate.
+In our case, the region is a single point, which has volume $0$.
+Therefore we have $k \log 0 = -\infty$: the entropy of a single microstate is minus infinity.
+Alternatively, we can use the Gibbs/Shannon entropy $ -\int \rho \log \rho \, dq^{3n} dp^{3n}$.\footnote{CITE} Our macrostate is described by a delta distribution centered on a single microstate: $\delta(q^i-q^i_0, p^i-p^i_0)$.
 The distribution is zero everywhere, except for the microstate where it is infinite.
 Therefore the integral returns minus infinity.
+No matter how we calculate it, the entropy of a single microstate is minus infinity.
 
-No matter our definition, then, the entropy of a single microstate is minus infinity. This certainly does not sound terribly reassuring... but is it a problem?
+The question now becomes: does it make physical sense for a state to have minus infinite entropy? If not, what are the consequences?
 
-First of all, this is in direct contradiction with the third law of thermodynamics.
-While the third law has many different formulations, it always leads to requiring the entropy of a perfect crystal at zero Kelvin to be zero.
-The idea is that a perfect crystal at zero Kelvin only has one possible configuration: it has one possible microstate.
-But we have seen that the entropy of a single microstate is minus infinity, not zero.
-Therefore the result is in contradiction.
+First of all, this is in direct contradiction to the third law of thermodynamics.
+While the third law has many different formulations, it always leads to requiring the entropy of a perfect crystal at zero Kelvin to be zero.\footnote{This is often taken to mean that a perfect crystal at zero Kelvin is frozen into a single configuration, but, as we saw before, that is not true.
+Zero entropy corresponds to a cell of unitary volume or to a uniform distribution over a unitary volume.
+Discrete and continuous spaces work differently at a measure theoretic level, which make both probability and entropy work differently as well.} Since the entropy for a single microstate is minus infinity, there is a clear violation.
 
@@
 If we transfer that increase of entropy to our minus infinite entropy system, we reset the entropy of the engine and the reservoirs to the initial state.
-Yet this leaves the entropy of our system unchanged: minus infinity plus a finite number is still minus infinity.
+Yet this leaves the entropy of our system unchanged: minus infinity plus a finite number is still minus infinity.\footnote{This is essentially a redressing of Maxwell's demon.
+Identifying any variable with infinite precision collapses the phase space volume of possible microstates to zero, giving a state that violates the third law and that would allow to violate the second as well.}
 
@@
 But, as we said, no such thing exists.
-So, from an information perspective, no system can have entropy lower than zero.\footnote{We believe this to be the true form of the third law.} A single classical microstate violates that.
+So, from a conceptual perspective, no system can have entropy lower than zero.\footnote{In reverse physics, we call this the Principle of Maximal description.} A single classical microstate violates that.
 
-So classical mechanics fails because it allows for the possibility of states that can never exist, those with minus infinite entropy.
+We are in the paradoxical position that classical mechanics is founded on states with perfectly defined continuous quantities that, on thermodynamic grounds, can never be realized.
 On practical grounds, we could bypass the second law with all its implications.
@@
 On logical grounds, they would provide more information about the system than stating that the system does not exist.
+But what is the physical cause of this contradiction?
 
-We want to stress here that these entropic arguments are not independent from the isolation arguments.
-Thermodynamics as a physical theory is exactly about describing what happens across the boundary.
-Heat and work are energy exchanges by one system to others.
-Perfect control of a system is effectively Maxwell’s demon in disguise.
+Let us look at it from an information theoretic sense.
+Suppose we use the state of a physical system to encode some data.
+That is, we choose the state based on the specific values we want to encode.
+And suppose we let the system reach equilibrium with the environment.
+The higher the entropy of the final state, the more information will be lost.
+In other words, we can understand the entropy as an indicator of the disturbance the system-environment interaction placed on the state of the system.
+Given that the state of a classical system is fully specified by continuous quantities, we would be able to encode an infinite amount of information on a finite system.
+Conversely, any finite amount of noise introduced by the environment on the system would mean the loss of infinite information.
+If we imagine that there is a minimum amount of system-environment interaction that cannot be avoided, then this sets the zero of how much information can be encoded or retrieved on a physical system.
 
-Those who are familiar with communication theory are aware that the channel capacity is given by the signal-to-noise ratio.\footnote{CITE: papers that define channel capacity (Shannon?)} In physics terms, the noise is the background interaction of the system with the environment.
-So that fixes the zero: we cannot use the system to encode more information than the background interaction allows.
-Saying that we can reduce the interaction with the system as much as we want effectively means setting the noise level to minus infinity, removing the background, perfectly isolating the system.
-In this light, all of these arguments are actually the same argument: we cannot fully isolate a system, we cannot encode infinite information in a state, we cannot have an infinite entropy differential.
+Before we explore more what happens across the system-environment boundary, we note that quantum mechanics does solve this problem.
+The entropy of any pure state in quantum mechanics is zero.
+The entropy of any mixed quantum state is greater then zero.
+Therefore there are no quantum states with negative entropy.
+In this light, we can picture the classical description becoming more and more untenable as entropy decreases.
+That is, in the same way that we understand non-relativistic mechanics as the limit of relativistic mechanics when velocities are much smaller than $c$, we should understand classical mechanics as the limit of quantum mechanics for states whose entropy is much greater than $0$.
 
-\section{Quantum resolutions and open problems}
+\section{System-environment boundary}
 
-While this paper is not about providing a clear and complete understanding of quantum mechanics, we want to give a sense of how the new theory partially solves some of the problems of the old one.
+Let us now proceed by studying the system-environment boundary.
+It should be clear that the very definition of a physical system implies the definition of a boundary.
+Whether we are studying a solid object, an open container of gas, a chemical compound or a single particle, we must declare what constitutes part of the system and what does not.
+Moreover, the nature of the interaction between the system and the environment at the boundary is critical for the very definition of state for the system.
+Let us start with the following simple example.
 
-To start, in quantum mechanics all pure states have zero entropy, including the vacuum, the ``empty'' system.
-Mixed states have positive entropy.
-We can no longer encode infinite information within the state of a system with continuous variables.
-Conceptually, the accessible information of the system is quantized by setting a lower limit, a zero, that cannot be crossed.
-This removes all the inconsistencies with thermodynamics, statistical mechanics and information theory.
-Conceptually, it signifies there is always a background interaction with the environment that cannot be fully eliminated.
+Suppose we want to study the motion of a cannon ball.
+As the cannon ball moves, air molecules will scatter off its surface.
+Yet, the effect produced will be negligible.
+We can therefore describe the state of the cannon ball with a single, infinitely precise, pair of position and momentum.
+Now suppose we want to study the motion of a speck of dust.
+This time the effect of the air molecules will not be negligible.
+We will therefore describe the state of the speck of dust with a probability distribution over position and momentum.
+Now suppose we want to study the motion of the cannon ball on the surface of the sun.
+Plasma will scatter off the ball catastrophically, making the notion of the state of ball meaningless.
 
-If there is interaction at the system/environment boundary, then the properties of the system can only be defined up to the statistical fluctuations imposed by the interaction.
-In fact, in quantum mechanics the probabilistic description is baked into the very notion of state, it is not something that can be removed like in classical statistical mechanics.
+The definition of state is contingent upon the system-boundary interaction.\footnote{We use ``system'' and ``environment'' as it mostly evoke the right intuition.
+However, the same considerations apply to the internal dynamics that is below the sensitivity of our instruments.
+For example, a proton is described as a single quantum system precisely at those energies for which the internal quark-gluon dynamics can be disregarded.
+More precisely, ``system'' includes only those properties that are under study given the chosen processes.} If we says that we are describing a cannonball with infinitely precise position and velocity, we are implicitly stating that the we are not at 5000 Kelvin on the surface of the sun.
+The most precise definition for the state of a system, then, is the one that can be given in the best conditions: when the system-boundary interaction causes the least disruption.
 
-If the system/environment interaction at the boundary participates in the definition of the state of a system, then states cannot be understood as intrinsic properties of the system that exist before a particular type of interaction has taken place.
-This is exactly what happens in quantum mechanics: an observable is well defined only when the system is put in determined conditions.
-A measurement may change those conditions as it is, in general, a change of the system/environment interaction.
-Contextuality is therefore a feature, not a problem to get rid of, and we should realize it is not even an exclusive feature of quantum mechanics.
+If the state of a system is to be described by infinitely precise quantities, this means that there exists at least one setting for which the system is completely isolated.
+But complete isolation does not make physical sense.
+In practice, this is not something that can be realized.
+But even theoretically possible it is not possible: any wall we will use as shielding will necessarily have finite temperature and will emit thermal radiation; gravitational interactions cannot be shielded in the first place.\footnote{CITE: something that goes more into the details as to why gravitational interactions cannot be shielded?} Moreover, even conceptually complete isolation is problematic.
+It would mean that no signal, electromagnetic, gravitational or otherwise, would be able to interact with the object.
+It would be effectively experimentally inaccessible, physically ill defined.
+This means that there must be a minimal unavoidable interaction between the system and the environment.
 
+If the uncertainty in our measurements or if other sources of noise are great compared with this minimal interaction, then the latter can be disregarded and the state of the system can be modeled as an infinitely precise quantity.
+The bigger uncertainty ``hides'' the complication of the minimal interaction.
+But if this minimal source is the dominant part, we cannot model it as some uncertainty over some infinitely precise value: it does not work conceptually.
+It would mean that, in line of principle, that minimal interaction could be removed, which, by definition, it cannot.\footnote{As an analogy, to the naked eye a star looks like a point in the sky and can be assumed to be as such.
+Even with several order of magnification, the assumption will hold.
+However, at some point one reaches a magnification such that the star becomes bigger and the assumption break down.
+At that point, you can't describe the star as just a single point: it's a more complicated object.
+Counter-intuitively, the infinitely precise value is the approximation.}
+
+Note that this maps perfectly with the previous entropic argument.
+The zero entropy limit corresponds to this minimal interaction.
+Classical mechanics works when the leading source of uncertainty is much greater than the minimal interaction, when the entropy of the states is much greater than zero.
+As we remove sources of noise and uncertainty, the minimal interaction becomes dominant, and classical mechanics fails.
+
+The problem with classical mechanics is not so much that it assumes perfect isolation, but that it assumes it at all level of scale.
+With continuous materials, we can take smaller and smaller parts.
+Classical fields allow for wave packets of ever decreasing energy.
+The theory implicitly assumes all these objects can be prepared, manipulated and detected independently enough that their state can be perfectly defined.
+This is the inherent problem with classical mechanics: perfect isolation itself is a simplifying assumption and cannot be taken to always hold.
+As we reach smaller and smaller scales, study smaller and smaller objects the assumption becomes less and less tenable.
+
+\section{Discussion}
+
+We would like to stress that the arguments we have provided are straight forward physical considerations.
+There is no mystery.
+They do not require infinite parallel worlds, special roles of observers, subjective information or many of the other esoteric elements present in various interpretations of quantum mechanics.
+Just simple considerations about thermodynamics and system-environment interaction.
+
+Under this light, there is a natural motivation that tells us why quantum mechanics is a statistical theory, and it is so at a fundamental level.
+What is ``quantized'' is the entropy due to an unavoidable background.\footnote{As discussed in a previous footnote, this may also include decoupling from the internal dynamics of the system.} Identifying a physical system means identifying that experimentally accessible description, that information, that decouples well from everything else, such that we can provide an independent description and rules of evolution.
+And that description is contingent to a particular system-environment interaction.
+
+An interaction with the system, then, may not only affect the state of the system, but the system-environment interaction.
+This is, again, nothing unexpected, as these are the type of situations thermodynamics and statistical mechanics describe.
 In fact, suppose we have an open flask of gas in equilibrium.
@@
 The number of particles $N$ of the gas will not be well-defined as it fluctuates, and it is described by a probability distribution.
-If we want a state where $N$ is well-defined, we can simply close the flask and let the system equilibrate.
-We will obtain a canonical ensemble determined by the same $T$ and $V$ and by the particle number $N$.
+If we want a state in which $N$ is well-defined, we can simply close the flask and let the system equilibrate.
+We will obtain a canonical ensemble determined by the same $T$ and $V$ and by the particle number $N$, and the chemical potential $\mu$ is undefined.
 The number of particles will depend on the fluctuation at the time we closed the flask, therefore we will have a certain probability to have a particular $N$.
@@
 Note the parallel with quantum mechanics: we start with a state where an observable, say the energy, is well defined and we end in a state where another observable, say position, is well defined with a value determined by a probability distribution.
-We change context, we change type of ensemble, we change what quantities define the ensemble.
-The point is that quantum contextuality is not the exception, it is classical \emph{non}-contextuality that is the exception.
 
-For these types of reasons quantum mechanics is more precise than classical mechanics: because it better captures the definition of our system.
-It captures the effect of a constant minimal interaction of the system  with the environment which classical mechanics simply disregards.
-If anything, quantum mechanics does not go far enough.
-
-For example, while the accessible information about the system is quantized, the possible values of the observables themselves are not.
-That is, position and momentum are still modeled by a real quantity.
-This implicitly assumes that our detectors can be made arbitrarily precise, that the resolution of our equipment can always be made finer than the resolution characterized by the quantum system.
-But our equipment, supposedly, is made of quantum systems.
-Therefore, at some point, this assumption has to fail.
-These types of metrological requirements will ultimately need to be incorporated into a more precise theory.
-
-Another issue is that all pure quantum states have zero entropy and therefore the entropy cannot change during evolution.
-This makes it impossible to characterize what happens during out-of-equilibrium transitions, when the behavior at the boundary drastically changes.
-Note that quantum projections are exactly of this type.
-The temptation is simply to include part of the environment within the system, and then trace it out.
-This assumes that the interaction between the system and the part of the environment is such that it can be in principle ``switched off'', as in the case of the air with the speck of dust, and therefore does not take part in the definition of the system itself.
-However, it is conceivable that, while out of equilibrium, the overall system temporarily reaches states where each component cannot be described as a separate independent object.
-This type of interaction would necessarily influence the definition of the system, and therefore cannot be simply layered on top of the individual components.
-This would seem out of reach of the current quantum framework.
+The development of a full fledged understanding of quantum mechanics goes beyond the scope of this article.
+We do want to show, however, that such a straight forward picture, based on tried and true old physical ideas, is possible.
+We feel that the overall physics community is ``stuck'' thinking that quantum mechanics is fundamentally mysterious, and therefore a new set of concepts are needed.
+This means that we will never find a down-to-earth, intuitive and obvious explanation.
+Because we can only find what we are looking for.
+We choose to believe such an explanation exists, and we just have to clean it up.
+We do that because, if we don't, we will never find one.
+We hope that, by showing that the failure of classical mechanics is, in retrospect, reasonable, we can find a ``reasonable interpretation'' of quantum mechanics.
 
@@
 
-We believe these arguments help shed light onto why classical mechanics fails at a conceptual level and pave the way to a better understanding of quantum mechanics.
-They also give insights into what other problems still need to be addressed at a fundamental level.
+We have shown that classical mechanics lies on the implicit assumption of perfect isolation at all scale: every fragment of matter, every minuscule wave packet, can be produced, manipulated and measured with infinitesimal precision.
+This is not only untenable, but it is also in conflict with thermodynamics.
+The existence of a minimal, unavoidable, interaction with the environment, with the uncertainty associated with it, is ultimately what makes the classical ideas fail.
 
-We close with the following general remark.
-Much of the work on the foundations of physics nowadays is geared toward finding the ``true'' objects and rules of nature.
-\footnote{CITE: papers that propose emergent models of space time, ...} Even in quantum mechanics, a lot of efforts are aimed at finding the true underlying constituents and mechanisms that would explain the quantum world.\footnote{CITE: papers that propose underlying mechanism (e.g.
-Bohmian approach, GRW)} In this sense, theoretical physics is built bottom-up.
-We believe this general approach to be problematic.
-
-Experimental physics is built top-down: we start with big objects, and decompose them into smaller and smaller parts.
-The objects of our research are defined by what parts can be carved out, accessed and manipulated experimentally, independently from one another.
-This imposes a number of requirements, constraints and assumptions that allow us to define the physical system.
-These definitional requirements are, therefore, not practical considerations that only experimentalists should be bothered by.
-They must be at the foundation of our theoretical treatment since, without their stipulation, there is no physical system to study.
-In this sense, they are more fundamental than the laws the system obeys, because we cannot define the law before having a well-defined system.
-
-We believe it is the inclusion of these definitional requirements and assumptions, which we have started to sketch here and in our more general project Assumptions of Physics, that will ultimately form a more sound foundation for physical theories.
+This is a definitional issue, as the very definition of a system impinges on finding some experimentally accessible properties that decouples from everything else.
+The insight that nothing can be fully decoupled, then, has to be included at a fundamental level in the definition of a system, and this is what quantum mechanics, albeit in a still unclear way, must be doing.
+We believe a full understanding of these definitional requirements and constitutive assumptions are needed to give a more sound foundation to all current physical theories, and to open the way for future ones.
 

### 2023-02-02 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
ClassicalFailure: reviewed with Christine

@@
 Therefore we have $k \log 0 = -\infty$: the entropy of a single microstate is minus infinity.
-Alternatively, we can use the Gibbs/Shannon entropy $ -\int \rho \log \rho \, dq^{3n} dp^{3n}$.\footnote{CITE} Our macrostate is described by a delta distribution centered on a single microstate: $\delta(q^i-q^i_0, p^i-p^i_0)$.
+Alternatively, we can use the Gibbs/Shannon entropy, $ -\int \rho \log \rho \, dq^{3n} dp^{3n}$.\footnote{CITE} Our macrostate is described by a delta distribution centered on a single microstate: $\delta(q^i-q^i_0, p^i-p^i_0)$.
 The distribution is zero everywhere, except for the microstate where it is infinite.
@@
 Yet this leaves the entropy of our system unchanged: minus infinity plus a finite number is still minus infinity.\footnote{This is essentially a redressing of Maxwell's demon.
-Identifying any variable with infinite precision collapses the phase space volume of possible microstates to zero, giving a state that violates the third law and that would allow to violate the second as well.}
+Identifying any variable with infinite precision collapses the phase space volume of possible microstates to zero, giving a state that violates the third law and that would allow one to violate the second as well.}
 
@@
 But, as we said, no such thing exists.
-So, from a conceptual perspective, no system can have entropy lower than zero.\footnote{In reverse physics, we call this the Principle of Maximal description.} A single classical microstate violates that.
+So, from a conceptual perspective, no system can have entropy lower than zero.\footnote{In Reverse Physics, we call this the Principle of Maximal Description.} A single classical microstate violates that.
 
-We are in the paradoxical position that classical mechanics is founded on states with perfectly defined continuous quantities that, on thermodynamic grounds, can never be realized.
+We are in the paradoxical position that classical mechanics is founded on states with perfectly defined continuous quantities that, thermodynamically, can never be realized.
 On practical grounds, we could bypass the second law with all its implications.
@@
 
-Let us look at it from an information theoretic sense.
+Let us look at it from an information theoretic perspective.
 Suppose we use the state of a physical system to encode some data.
@@
 Conversely, any finite amount of noise introduced by the environment on the system would mean the loss of infinite information.
-If we imagine that there is a minimum amount of system-environment interaction that cannot be avoided, then this sets the zero of how much information can be encoded or retrieved on a physical system.
+If we imagine that there is a minimum amount of system-environment interaction that cannot be avoided, then this sets the zero of how much information can be encoded into the state of a physical system without loss.
+Therefore we can understand this minimum amount of system-environment interaction as the physical cause of the zero entropy limit.
 
-Before we explore more what happens across the system-environment boundary, we note that quantum mechanics does solve this problem.
-The entropy of any pure state in quantum mechanics is zero.
-The entropy of any mixed quantum state is greater then zero.
+Before we explore more what happens across the system-environment boundary, we note that quantum mechanics does suffer from the same problem.
+The entropy of any pure state in quantum mechanics is zero, not minus infinity.
+The entropy of any mixed quantum state is greater than zero.
 Therefore there are no quantum states with negative entropy.
@@
 
-The definition of state is contingent upon the system-boundary interaction.\footnote{We use ``system'' and ``environment'' as it mostly evoke the right intuition.
+The definition of state is contingent upon the system-boundary interaction.\footnote{We use ``system'' and ``environment'' as these terms mostly evoke the right intuition.
 However, the same considerations apply to the internal dynamics that is below the sensitivity of our instruments.
 For example, a proton is described as a single quantum system precisely at those energies for which the internal quark-gluon dynamics can be disregarded.
-More precisely, ``system'' includes only those properties that are under study given the chosen processes.} If we says that we are describing a cannonball with infinitely precise position and velocity, we are implicitly stating that the we are not at 5000 Kelvin on the surface of the sun.
+More precisely, ``system'' includes only those properties that are under study given the chosen processes.} If we say that we are describing a cannon ball with infinitely precise position and momentum, we are implicitly stating that the we are not at 5000 Kelvin on the surface of the sun.
 The most precise definition for the state of a system, then, is the one that can be given in the best conditions: when the system-boundary interaction causes the least disruption.
@@
 In practice, this is not something that can be realized.
-But even theoretically possible it is not possible: any wall we will use as shielding will necessarily have finite temperature and will emit thermal radiation; gravitational interactions cannot be shielded in the first place.\footnote{CITE: something that goes more into the details as to why gravitational interactions cannot be shielded?} Moreover, even conceptually complete isolation is problematic.
+But even theoretically it is not possible: any wall we will use as shielding will necessarily have finite temperature and will emit thermal radiation; gravitational interactions cannot be shielded in the first place.\footnote{CITE: something that goes more into the details as to why gravitational interactions cannot be shielded?} Moreover, even conceptually complete isolation is problematic.
 It would mean that no signal, electromagnetic, gravitational or otherwise, would be able to interact with the object.
@@
 It would mean that, in line of principle, that minimal interaction could be removed, which, by definition, it cannot.\footnote{As an analogy, to the naked eye a star looks like a point in the sky and can be assumed to be as such.
-Even with several order of magnification, the assumption will hold.
-However, at some point one reaches a magnification such that the star becomes bigger and the assumption break down.
+Even with several factors of magnification, the assumption will hold.
+However, at some point one reaches a magnification such that the assumption breaks down.
 At that point, you can't describe the star as just a single point: it's a more complicated object.
@@
 The zero entropy limit corresponds to this minimal interaction.
-Classical mechanics works when the leading source of uncertainty is much greater than the minimal interaction, when the entropy of the states is much greater than zero.
-As we remove sources of noise and uncertainty, the minimal interaction becomes dominant, and classical mechanics fails.
+Classical mechanics works when the leading source of uncertainty is much greater than the minimal interaction, when the entropy of the state is much greater than zero.
+As we remove sources of noise and uncertainty, eventually the minimal interaction becomes dominant, and classical mechanics fails.
 
-The problem with classical mechanics is not so much that it assumes perfect isolation, but that it assumes it at all level of scale.
-With continuous materials, we can take smaller and smaller parts.
+The problem with classical mechanics is not so much that it assumes perfect isolation, but that it assumes it at all scales.
+Classical fluids and continuous materials allow us to study smaller and smaller parts.
 Classical fields allow for wave packets of ever decreasing energy.
 The theory implicitly assumes all these objects can be prepared, manipulated and detected independently enough that their state can be perfectly defined.
-This is the inherent problem with classical mechanics: perfect isolation itself is a simplifying assumption and cannot be taken to always hold.
-As we reach smaller and smaller scales, study smaller and smaller objects the assumption becomes less and less tenable.
+This is the inherent problem with classical mechanics: perfect isolation at all scales.
+In the correct circumstances, perfect isolation is an internally consistent assumption.
+In the correct circumstances, recursive subdivision into smaller parts is an internally consistent assumption.
+It is the joint assumption that the system can be recursively subdivided into perfectly isolated parts that is physically untenable.\footnote{Note that the mathematical model is still perfectly consistent.
+It is the physical model that is not.
+That is why one should never confuse mathematical consistency with physical consistency.} As we consider smaller and smaller scales, study smaller and smaller objects at an ever increasing level of precision, we are inevitably bound to move away from the regime where perfect isolation can be assumed.
 
@@
 
-We would like to stress that the arguments we have provided are straight forward physical considerations.
+We would like to stress that the arguments we have provided are straightforward physical considerations.
 There is no mystery.
@@
 
-Under this light, there is a natural motivation that tells us why quantum mechanics is a statistical theory, and it is so at a fundamental level.
+In light of this, there is a natural motivation as to why quantum mechanics is a statistical theory, and why it is so at a fundamental level: it incorporates the effect of a minimal interaction with the environment.
 What is ``quantized'' is the entropy due to an unavoidable background.\footnote{As discussed in a previous footnote, this may also include decoupling from the internal dynamics of the system.} Identifying a physical system means identifying that experimentally accessible description, that information, that decouples well from everything else, such that we can provide an independent description and rules of evolution.
-And that description is contingent to a particular system-environment interaction.
+And that description is contingent to a particular system-environment interaction: it is contextual.
 
 An interaction with the system, then, may not only affect the state of the system, but the system-environment interaction.
-This is, again, nothing unexpected, as these are the type of situations thermodynamics and statistical mechanics describe.
+This is, again, nothing mysterious, as these are the types of situations thermodynamics and statistical mechanics describe.
 In fact, suppose we have an open flask of gas in equilibrium.
@@
 If we want a state in which $N$ is well-defined, we can simply close the flask and let the system equilibrate.
-We will obtain a canonical ensemble determined by the same $T$ and $V$ and by the particle number $N$, and the chemical potential $\mu$ is undefined.
+We will obtain a canonical ensemble determined by the same $T$ and $V$ and by the particle number $N$, and the chemical potential $\mu$ is now undefined.
 The number of particles will depend on the fluctuation at the time we closed the flask, therefore we will have a certain probability to have a particular $N$.
-In short, we start with a state where $\mu$ is well-defined but not $N$, we interact with the system and end up in a state where $N$ is well defined and its value determined by a probability distribution.
+In short, we start with a state where $\mu$ is well-defined but not $N$; we interact with the system and end up in a state where $N$ is well-defined and its value determined by a probability distribution.
 Note the parallel with quantum mechanics: we start with a state where an observable, say the energy, is well defined and we end in a state where another observable, say position, is well defined with a value determined by a probability distribution.
@@
 The development of a full fledged understanding of quantum mechanics goes beyond the scope of this article.
-We do want to show, however, that such a straight forward picture, based on tried and true old physical ideas, is possible.
-We feel that the overall physics community is ``stuck'' thinking that quantum mechanics is fundamentally mysterious, and therefore a new set of concepts are needed.
-This means that we will never find a down-to-earth, intuitive and obvious explanation.
-Because we can only find what we are looking for.
-We choose to believe such an explanation exists, and we just have to clean it up.
-We do that because, if we don't, we will never find one.
-We hope that, by showing that the failure of classical mechanics is, in retrospect, reasonable, we can find a ``reasonable interpretation'' of quantum mechanics.
+We do want to show, however, that such a straightforward picture, based on tried and true old physical ideas, is possible.
+We feel that the overall physics community is ``stuck'' thinking that quantum mechanics is fundamentally mysterious, and therefore a new set of concepts is needed.
+The issue is that one can only find what one is looking for.
+If we want a down-to-earth, intuitive and obvious explanation, we must believe such a thing exists, or we will never find one.
+And if that's the case, what's the point?
+We hope that, by showing that the failure of classical mechanics is, in retrospect, reasonable, even necessary, we can find a ``reasonable interpretation'' of quantum mechanics.
 
@@
 
-We have shown that classical mechanics lies on the implicit assumption of perfect isolation at all scale: every fragment of matter, every minuscule wave packet, can be produced, manipulated and measured with infinitesimal precision.
+We have shown that classical mechanics lies on the implicit assumption of perfect isolation at all scales: every fragment of matter, every minuscule wave packet, can be produced, manipulated and measured with infinitesimal precision.
 This is not only untenable, but it is also in conflict with thermodynamics.
-The existence of a minimal, unavoidable, interaction with the environment, with the uncertainty associated with it, is ultimately what makes the classical ideas fail.
+The existence of an unavoidable minimal interaction with the environment, with the uncertainty associated with it, is ultimately what makes the classical ideas fail.
 
-This is a definitional issue, as the very definition of a system impinges on finding some experimentally accessible properties that decouples from everything else.
+This is a definitional issue, as the very definition of a system impinges on finding some experimentally accessible properties that decouple from everything else.
 The insight that nothing can be fully decoupled, then, has to be included at a fundamental level in the definition of a system, and this is what quantum mechanics, albeit in a still unclear way, must be doing.

### 2023-02-03 Notes/2023-QuantumGeneralizedCoordinates/QuantumGeneralizedCoordinates.tex
Making explicit the derivation of covariance of momentum
Updated from discussion with Lorenzo

@@
 		d\hat{x}^i &= \partial_j \hat{x}^i dx^j \\
-		d\hat{x}^n &= | \partial_j \hat{x}^i | dx^n = |J| dx^n \\
+		d\hat{x}^1 \cdots d\hat{x}^n &= | \partial_j \hat{x}^i | dx^1 \cdots dx^n = |J| dx^1 \cdots dx^n \\
 	\end{aligned}
@@
 	\begin{aligned}
-		\int \rho(x^j) dx^n &= \int | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i) dx^n = \int \hat{\rho}(\hat{x}^i) d\hat{x}^n \\
+		\int \rho(x^j) dx^1 \cdots dx^n &= \int | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i) dx^1 \cdots dx^n = \int \hat{\rho}(\hat{x}^i) d\hat{x}^1 \cdots d\hat{x}^n \\
 		|J| \hat{\rho}(\hat{x}^i) &= | \partial_j \hat{x}^i | \hat{\rho}(\hat{x}^i)  = \rho(x^j) \\
@@
 
+We can rediscover the transformation rules by expanding the state in a different basis.
+\begin{equation}
+	\begin{aligned}
+		\< \psi | \psi \> &= \< \psi | \int dx | x \> \< x | \psi\>  = \< \psi | \int d\hat{x} | \hat{x} \> \< \hat{x} | \psi\> \\
+		&= \< \psi | \int ||J|| dx | \hat{x} \> \hat{\psi}(\hat{x}) \\
+		&= \< \psi | \int dx \psi(x) \sqrt{||J||} | \hat{x} \>  \\
+		| x \> &= \sqrt{||J||} | \hat{x} \>
+	\end{aligned}
+\end{equation}
+Note that $|\psi\> = |\hat{\psi}\>$ because the state is coordinate independent, in the sense that it doesn't change depending in which coordinates is expressed.
+The delta functions $| x \>$ associated with position, however, do not work in the same way.
+As these are elements of a continuous spectra, and therefore are associated with densities, we have $| x \>$ $| x \> \neq | \hat{x} \>$ as they depend on the coordinate explicitly (i.e.
+the normalization constant changes).
+This would not happen, for example, if the spectra is discrete, like for the eigenstate of a harmonic oscillator.
+
 \section{Change of coordinates for momentum}
@@
 They are derived by making an arbitrary change on the position coordinates and inducing a coordinate change on momentum such that the symplectic form $\omega = dx^i \wedge dp_i$ is invariant.
-Therefore this depends on the symplectic structure.
-This makes conjugate momentum vary as a covector.
+That is:
+\begin{equation}
+	\begin{aligned}
+		\omega &= d\hat{x}^i \wedge d\hat{p}_i = \partial_j \hat{x}^i dx^j \wedge d\hat{p}_i = dx^j \wedge dp_j \\ 
+		d p_j &= \partial_j \hat{x}^i d\hat{p}_i
+	\end{aligned}
+\end{equation}
+
+Therefore this depends on the symplectic structure. This makes conjugate momentum vary as a covector.
 
@@
 		d \hat{p}_i &= \partial_i x^j d p_j \\
-		d\hat{p}^n &= | \partial_i x^i | dp^n = |J^{-1}| dx^n \\
+		d\hat{p}^n &= | \partial_i x^j | dp^n = |J^{-1}| dx^1 \cdots dx^n \\
 	\end{aligned}
@@
 	\begin{aligned}
-		\phi(p_j) &= F(\psi(x^i)) = \frac{1}{\sqrt{2 \pi \hbar}} \int e^{\frac{p_j x^j}{\imath \hbar}} \psi(x^j) dx^n \\
+		\phi(p_j) &= F(\psi(x^i)) = \frac{1}{\sqrt{2 \pi \hbar}} \int e^{\frac{p_j x^j}{\imath \hbar}} \psi(x^j) dx^1 \cdots dx^n \\
 	\end{aligned}
@@
 
+\section{Examples}
+
+\subsection{2D Cartesian and 2D polar}
+
+In Cartesian coordinates, the eigenstates of momentum correspond to oscillations at a fixed frequency along the corresponding axis.
+\begin{equation}
+	\begin{aligned}
+		P_x &= - \imath \hbar \partial_x \\
+		\phi(x,y) &= e^{\imath k_x x} \\
+		P_x e^{\imath k_x x} &= - \imath \hbar \partial_x e^{\imath k_x x} = - \imath \hbar \imath k_x e^{\imath k_x x} = \hbar k_x e^{\imath k_x x} \\ 
+		P_y &= - \imath \hbar \partial_y \\
+		\phi(x,y) &= e^{\imath k_y y} \\
+		P_y e^{\imath k_y y} &= - \imath \hbar \partial_y e^{\imath k_y y} = - \imath \hbar \imath k_y e^{\imath k_y y} = \hbar k_y e^{\imath k_y y} \\ 
+	\end{aligned}
+\end{equation}
+
+The Fourier transform, then, corresponds to
+\begin{equation}
+	\begin{aligned}
+		\phi(p_x, p_y) &= \frac{1}{\sqrt{2 \pi \hbar}} \int \frac{1}{\sqrt{2 \pi \hbar}} \int e^{\frac{p_x x}{\imath \hbar}} e^{\frac{p_y y}{\imath \hbar}} \psi(x, y) dxdy \\
+	\end{aligned}
+\end{equation}
+where the factors in the transform pick up the oscillations at the appropriate frequency.
+
+In polar coordinates, the eigenstates of momentum in $r$ and $\phi$ are different.
+For $r$, the oscillation will be along the radial axis (intuitively, like the ripples created by an object falling on water) while for $\phi$ the oscillation will depend on the angle.
+Eigenstates of momentum for $\phi$ will have to be periodic, and therefore the spectra will be discrete.
+\begin{equation}
+	\begin{aligned}
+		P_r &= - \imath \hbar \partial_r \\
+		\phi(r,\phi) &= e^{\imath k_r r} \\
+		P_r e^{\imath k_r r} &= - \imath \hbar \partial_r e^{\imath k_r r} = - \imath \hbar \imath k_r e^{\imath k_r r} = \hbar k_r e^{\imath k_r r} \\ 
+		P_\phi &= - \imath \hbar \partial_\phi \\
+		\phi(r,\phi) &= e^{\imath \frac{ n \phi}{2 \pi}} \\
+		P_\phi e^{\imath \frac{ n \phi}{2 \pi}} &= - \imath \hbar \partial_\phi e^{\imath \frac{ n \phi}{2 \pi}} = - \imath \hbar \imath \frac{ n }{2 \pi} e^{\imath \frac{ n \phi}{2 \pi}} = \frac{ \hbar n }{2 \pi} e^{\imath \frac{ n \phi}{2 \pi}r} \\
+	\end{aligned}
+\end{equation}
+The Fourier transform in polar coordinates is
+\begin{equation}
+	\begin{aligned}
+		\phi(p_r, p_\phi) &= \frac{1}{\sqrt{2 \pi \hbar}} \int \frac{1}{\sqrt{2 \pi \hbar}} \int e^{\frac{p_r r}{\imath \hbar}} e^{\frac{p_\phi \phi}{\imath \hbar}} \psi(r, \phi) dr d\phi \\
+	\end{aligned}
+\end{equation}
+so that the factors in the transform correctlypick the oscillations on the variables.
+
+Let's double check what happens with coordinate changes.
+\begin{equation}
+	\begin{aligned}
+		x &= r \cos \phi \\
+		y &= r \sin \phi \\
+		\partial_r x &= \cos \phi \\
+		\partial_\phi x &= - r \sin \phi \\
+		\partial_r y &= \sin \phi \\
+		\partial_\phi y &= r \cos \phi \\
+	\end{aligned}
+\end{equation}
+
+Express both momentum and eigenstate in other coordinates and see that it works.
+\begin{equation}
+	\begin{aligned}
+		P_r &= \partial_r x P_x + \partial_r y P_y = \cos \phi P_x + \sin \phi P_y = \frac{x}{\sqrt{x^2 + y^2}} P_x + \frac{y}{\sqrt{x^2 + y^2}} P_y  \\
+		e^{\imath k_r r} &= e^{\imath k_r \sqrt{x^2 + y^2}}\\
+		\partial_x e^{\imath k_r r} &= \partial_x e^{\imath k_r \sqrt{x^2 + y^2}} = e^{\imath k_r \sqrt{x^2 + y^2}} \imath k_r \partial_x \sqrt{x^2 + y^2} = \imath k_r \frac{x}{\sqrt{x^2 + y^2}} e^{\imath k_r \sqrt{x^2 + y^2}} \\
+		\partial_y e^{\imath k_r r} &= \partial_y e^{\imath k_r \sqrt{x^2 + y^2}} = e^{\imath k_r \sqrt{x^2 + y^2}} \imath k_r \partial_y \sqrt{x^2 + y^2} = \imath k_r \frac{y}{\sqrt{x^2 + y^2}} e^{\imath k_r \sqrt{x^2 + y^2}} \\
+		P_r e^{\imath k_r \sqrt{x^2 + y^2}} &= (-\imath \hbar ) \frac{x^2}{x^2 + y^2} \imath k_r e^{\imath k_r \sqrt{x^2 + y^2}} +(-\imath \hbar ) \frac{y^2}{x^2 + y^2} \imath k_r e^{\imath k_r \sqrt{x^2 + y^2}} \\
+		&= \hbar k_r \frac{x^2 + y^2}{x^2 + y^2} e^{\imath k_r \sqrt{x^2 + y^2}} = \hbar k_r e^{\imath k_r \sqrt{x^2 + y^2}} \\
+	\end{aligned}
+\end{equation}
+Note: we haven't transformed the wave function like a density.
+
+Conjecture: does the transform generalize like so:
+\begin{equation}
+	\begin{aligned}
+		\phi(p_{\hat{x}}, p_{\hat{y}}) &= \frac{1}{\sqrt{2 \pi \hbar}} \int \frac{1}{\sqrt{2 \pi \hbar}} \int e^{\frac{p_{\hat{x}} \hat{x}}{\imath \hbar}} e^{\frac{p_{\hat{y}} \hat{y}}{\imath \hbar}} \psi(\hat{x}, \hat{y}) d\hat{x}d\hat{y} \\
+	\end{aligned}
+\end{equation}
+
 \end{document}

### 2023-02-03 Articles/2023-PhysicalDifferentialTopology/PhysicalDifferentialTopology.tex
PhysicalDiffTopology: clarifying the problem

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amsfonts}

\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\begin{document}

\title{A physical foundation of Differential Topology}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}

\end{abstract}

\maketitle

\section{Introduction}

\section{Current problems}

The purpose of this work is to address a mismatch between the mathematical definitions of differential topology and the physical objects they need to represent. Both notationally and conceptually, these two worlds have drifted apart enough that the mathematical intuition that one needs to develop does not have a tight correspondence to the physical intuition. While one can make the translation, this is often too costly, leaves pieces behind and it severely lowers the availability and effectiveness of the tools. The goal is to find rigorous mathematical definition that are more natural for physics, make the physical assumptions explicit and generalize better. Here we list a series of specific problems we aim to address.

In differential topology, one defines local objects (i.e. vectors and forms) and then defines finite quantities associated to regions through integration. For example, one defines volume forms and then defines its integral over a region. Physics proceeds the opposite way. Experimentally, we do not measure directly a mass density, but rather we measure mass and volume of a finite region and take the ratio. We then imagine to take smaller and smaller regions, and the mass density is ratio in the limit. We want definitions that mirror this process so that we can understand more clearly what physical requirements and idealizations are embedded in the definitions. We find that it is a lot easier to build physical intuition on the finite objects than on the local ones.

Differentiabilty of a manifold is defined by requiring that the coordinate transformation among all charts are differentiable. All other notions, such as vectors as directional derivations or differentials as linear map, are therefore built on top. Therefore, we conceptually have two layer of definitions, one for ``analytical'' differentiability and one for ``topological'' differentiability. We would like to have a single unified notion that work directly at both levels. The hope is that the same definition can be generalized to the infinite dimensional case, so that we have a smoother transition to field theories.

While units and physical dimensions play a crucial role in physics, these ignored by the mathematics with important consequences. In differential geometry, vectors are defined as directional derivatives. But to define a derivative, we need to have a quantity to derive by. That quantity will have a unit. Therefore vectors are necessarily need a unit definition which is not part of the differentiable manifold, which is absolutely undesirable. For example, suppose $M$ is a two dimensional manifold, and suppose we use polar coordinates $[r, \theta]$. In differential topology, the directional derivative $\partial_r$  and $\partial_\theta$ represent a basis of the tangent space. But $r$ and $\theta$ have different physical dimensions. If $r$ is expressed in meters and $\theta$ is expressed in radians, $\partial_r$ has units of inverse meters and $\partial_\theta$ has units of inverse radians. The sum $\partial_r + \partial_\theta$ is ill-defined. The goal is to have core definitions that work naturally with units, and explicit rules that clarify units for all components of all objects.

\section{Differentials as sequences}

\section{Differentiability as reducibility}

\section{Tensors as maps between differentials}

\section{Reducible functionals and forms}

\bibliography{bibliography}

\end{document}

### 2023-02-06 Articles/2023-PhysicalDifferentialTopology/PhysicalDifferentialTopology.tex
PhysicalDiffTopology: reviewed motivation

@@
 
-The purpose of this work is to address a mismatch between the mathematical definitions of differential topology and the physical objects they need to represent.
-Both notationally and conceptually, these two worlds have drifted apart enough that the mathematical intuition that one needs to develop does not have a tight correspondence to the physical intuition.
-While one can make the translation, this is often too costly, leaves pieces behind and it severely lowers the availability and effectiveness of the tools.
-The goal is to find rigorous mathematical definition that are more natural for physics, make the physical assumptions explicit and generalize better.
-Here we list a series of specific problems we aim to address.
+Before proposing an alternative set of core definitions for differential topology, we should discuss what are the problems with the current ones.
+As with any choice, there is a certain element of personal preference, which we acknowledge.
+However, given that physics and mathematics have different goals, it should not be too surprising that they have different requirements.
+In particular, in physics formal objects have to model the real world, therefore definitions that map better to physical entities and make explicit the idealizations assumed by the model are objectively preferable.
+Such tools foster a mathematical intuition connected to the physical intuition, and considerably lowers the cognitive load.
+Let us explore, then, some areas in which the current tools are less than optimal.
 
-In differential topology, one defines local objects (i.e.
+In differential topology, one defines local objects (e.g.
 vectors and forms) and then defines finite quantities associated to regions through integration.
@@
 Experimentally, we do not measure directly a mass density, but rather we measure mass and volume of a finite region and take the ratio.
-We then imagine to take smaller and smaller regions, and the mass density is ratio in the limit.
-We want definitions that mirror this process so that we can understand more clearly what physical requirements and idealizations are embedded in the definitions.
-We find that it is a lot easier to build physical intuition on the finite objects than on the local ones.
+We then imagine to take smaller and smaller regions, and the mass density is ratio in the limit.\footnote{The Radon–Nikodym derivative, for example, follows this approach.} Ideally, our definitions should mirror this process to which there are at least to major advantages.
+First, it is a lot easier to build physical intuition on the finite objects than on the local ones.
+Second, we can understand more clearly what physical requirements and idealizations are embedded in the definitions.
+For example, one basic assumption to define the mass density is that if we partition the volume, the total mass is the sum of the one found in each part.
+Given the equivalence of mass and energy, this is strictly true only if there is no interaction energy between the parts.
+Exactly what happens is those cases, then, is not clear, and starting directly with local objects may hide actual conceptual problems.
 
-Differentiabilty of a manifold is defined by requiring that the coordinate transformation among all charts are differentiable.
+Differentiability of a manifold is defined by requiring that the coordinate transformation among all charts are differentiable.
 All other notions, such as vectors as directional derivations or differentials as linear map, are therefore built on top.
-Therefore, we conceptually have two layer of definitions, one for ``analytical'' differentiability and one for ``topological'' differentiability.
+Therefore, we conceptually have two layers of definitions, one for ``analytical'' differentiability and one for ``topological'' differentiability.
 We would like to have a single unified notion that work directly at both levels.
-The hope is that the same definition can be generalized to the infinite dimensional case, so that we have a smoother transition to field theories.
+Ideally, the same definition should be generalizable to the infinite dimensional case, so it could be directly applied to field theories.
 
 While units and physical dimensions play a crucial role in physics, these ignored by the mathematics with important consequences.
-In differential geometry, vectors are defined as directional derivatives.
-But to define a derivative, we need to have a quantity to derive by.
-That quantity will have a unit.
-Therefore vectors are necessarily need a unit definition which is not part of the differentiable manifold, which is absolutely undesirable.
-For example, suppose $M$ is a two dimensional manifold, and suppose we use polar coordinates $[r, \theta]$.
-In differential topology, the directional derivative $\partial_r$  and $\partial_\theta$ represent a basis of the tangent space.
-But $r$ and $\theta$ have different physical dimensions.
-If $r$ is expressed in meters and $\theta$ is expressed in radians, $\partial_r$ has units of inverse meters and $\partial_\theta$ has units of inverse radians.
+First of all, most people, physicists and mathematicians alike, miss that transformation rules of tensor components (i.e.
+covariance and contravariance) are primarily about units.
+That is, some quantities are expressed in base units (e.g.
+position and velocity) and others in derived units (e.g.
+velocity), therefore a change of definition in the first induces a change in the others.
+Since mathematics does not explicitly formalize units, it can do two things: either just define objects based on their transformation rules or look for objects that, by definition, follow the same transformation rules.\footnote{It seems that,in the mathematical world, the latter approach has been gaining favor with respect to the second.} To give an example of why this is problematic, in differential geometry one talks about tangent and cotangent bundle.
+Roughly speaking, the first pairs a point with a vector and the second pairs a point with a one form.
+It is often said that Lagrangian mechanics works on the tangent bundle, while Hamiltonian mechanics work on the cotangent bundle since one identifies velocity with a vector and momentum with a one form.
+Conceptually, however, the space in both cases represents all possible states of point particles.
+It is a manifold and the state is identified by a set of six variables.
+The fact that we choose position/velocity or position/momentum does not change the nature of the state space.
+Moreover, nothing prevents us to use units of velocity that are different from the one derived from the units of position.\footnote{This is actually often done for polar and cylindrical coordinates.
+The basis induced by the coordinates is not orthonormal, and is often replaced by one that is.
+It is not rare for people to not be aware of this and getting confused.} Therefore the notion of tangent/cotangent bundle is there to capture unit relationships between state variables.
+We can now ask: are there similar relationships that we may want to capture?
+If we have multiple particles, we will typically want to express their position in the same units.
+If we have the state space for $N$ particles, then, the transformation rules would be more restrictive than those of a tangent or cotangent bundle.
+
+The issue of units is even more crucial, as it breaks the notion of tangent space as it is typically defined in differential topology.
+There are two standard ways to define the tangent space: either as a variation of the point under a change of coordinates (i.e.
+the basis becomes $e_i = \partial_i P$) or as a directional derivative of a scalar function (i.e.
+the basis becomes $e_i = \partial_i$).
+Both use the notion of a derivative, which requires a parameter that, ultimately, defines the speed of the change.
+Physically, that parameters will have a unit.
+This presents two problems.
+First, the unit is an extra choice that cannot be uniquely defined by the choice of manifold.
+Second, different coordinates can be expressed as different units and, since we can only sum quantities with the same units, we would have a linear space for each choice of units/physical dimension.
+For example, suppose $M$ is a two dimensional manifold, and suppose we use polar coordinates $[r, \theta]$ with $r$ is expressed in meters and $\theta$ is expressed in radians.
+Let the directional derivatives $\partial_r$ and $\partial_\theta$ represent a basis of the tangent space.
+The former has units of inverse meters while the latter has units of inverse radians.
 The sum $\partial_r + \partial_\theta$ is ill-defined.
-The goal is to have core definitions that work naturally with units, and explicit rules that clarify units for all components of all objects.
 

### 2023-02-08 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HS: new version

--- 2022-01-27 Articles/2021-HarriganSpekkens/HarriganSpekkens.tex
@@
 \usepackage{hyperref}
+\usepackage{tikz}
 \hypersetup{colorlinks= true, allcolors=blue}
@@
 
-\section{Epistemic structure}
+\section{Entropy and the ontological model}
 
-Let us consider more closely the space of epistemic states given by $p(\lambda | P)$ and equation \ref{epistemic_mixing}.
-To make the discussion more concrete, we assume we have a $\psi$-complete model of a single qubit representing the direction of spin.
+To reproduce quantum theory, an ontological model has to reproduce not simply the probability of transitions during measurements but also all the results of quantum statistical mechanics and quantum information theory.
+Therefore it has to reproduce the correct values for entropy of mixed states.
+The natural way is by using standard information theoretic tools over the space of epistemic states which, as we will see, rules out $\psi$-ontic models.
+The degree of overlap between distributions, in fact, is not a choice, but is fixed by the value of entropy.
 
-Suppose we have a protocol that prepares either $z^+$ or $z^-$ with equal chance.
-This results in an epistemic state with equal distribution on $z^+$ and $z^-$.
-Similarly, suppose we have another protocol that prepares $x^+$ and $x^-$ with equal chance generating an epistemic state with equal distribution on $x^+$ and $x^-$.
-Note that such epistemic distributions do not overlap.
-Therefore, according to HS categorization, these correspond to two ontologically distinct states, making the preparation axis an ontological property \emph{of the system}, regardless of the orientation.
-However, the respective density matrices would not just overlap: they would be the same maximally mixed state.
-Hence there seems to be a mismatch between the type of epistemic knowledge one can express in QM through mixed states and the type of epistemic knowledge described in the HS model.
-This needs to be investigated further.
+Let us first review a crucial property of the information entropy, both in the classical setting (i.e.\ the Shannon/Gibbs entropy) and in the quantum setting (i.e.\ the von Neumann entropy).\footnote{We show the explicit calculations in the appendix.}
 
-Let us first characterize the extent of the discrepancy.
-In HS, the space of epistemic states $E_{HS}$ is the set of probability distributions over pure states, which is an infinite dimensional space.
-This is a direct consequence of defining the epistemic states as a distribution $p(\lambda|P)$ that, for each preparation $P$, assigns a well defined probability to each ontological state $\lambda$ \footnote{We will follow the convention of not distinguishing between probability and probability density.}.
-In QM, mixed states $E_{QM}$ are in one-to-one correspondence with the interior of the Bloch sphere, a three dimensional manifold.
-Looking at the difference in dimension, the discrepancy is not minor.
+\textbf{Classical information entropy of mixed non-overlapping distributions}.
+Let $\rho_1$ and $\rho_2$ be two classical distributions over a space $\Lambda$ with measure $\lambda$.
+The entropy for each distribution is given by the usual formula\footnote{The logarithm is assumed to be in base 2.}, for example
+\begin{equation}\label{shannon_entropy}
+	H(\rho_1) = - \int_\Lambda \rho_1(\lambda) \log \rho_1(\lambda) d\lambda.
+\end{equation}
+Suppose the two distributions are disjoint, and let $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ be a uniform mixture of the two distributions. The entropy of $\rho$ is given by
+\begin{equation}\label{entropy_nonoverlap}
+	H(\rho) = 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2).
+\end{equation}
+Note how the non-overlapping assumption fixes the entropy of the mixed state.
 
-Note that the space of pure states $\Lambda$ is the same in both cases, one state for each direction of spin.
-Also note that all epistemic states in both spaces can be reached by mixture of pure states.
-The difference is that, in the quantum case, we have an ``aliasing'': different mixtures will give the same mixed states.
-Since each classical probability distribution corresponds to a unique mixing of pure states, we have a map $\iota : E_{HS} \mapsto E_{QM}$ that allows us to go from HS-epistemic states to density matrices, the QM-epistemic states.
-This map is surjective but not injective.
-Therefore $E_{QM}$ can be seen as a set of equivalence classes of $E_{HS}$.
-The key question, then, is the following: is the map $\iota$ sufficiently well behaved that we can understand the epistemic content of $E_{QM}$ based on the epistemic content of $E_{HS}$?
-Or is it so ill behaved that there is no connection between the two?
+\textbf{Quantum information entropy of quantum mixed states}.
+Now suppose $\psi$ and $\phi$ are two pure quantum states and let $p = | \langle \psi | \phi \rangle |^2$ be the probability of transition from one to the other.
+Consider the mixed state $\rho = \frac{1}{2} | \psi \rangle \langle \psi | + \frac{1}{2} | \phi \rangle \langle \phi |$.
+Its entropy is given by
+\begin{equation}\label{entropy_mixed}
+	H(\rho) = H\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right).
+\end{equation}
 
-Ideally, one would like to say that $E_{HS}$ represents all possible epistemic states that include the knowledge about preparation, while $E_{QM}$ represents only those distinguishable through measurement.
-For example, one may want to say that the axis of preparation is, in principle, an ontologically distinct property, but this is not experimentally distinguishable.
-Therefore, distinct HS-epistemic states for each direction are projected by $\iota$ onto the same QM-epistemic state of maximum uncertainty.
-This sounds very plausible and appealing, but it turns out this is not the case.
-To show this fact we will concentrate on a single aspect of the map: how information entropy transforms under $\iota$.
-While there are other problems, we do not need to explore them: given the crucial role of entropy in information theory, its breakdown is sufficient to show that the epistemic structures of $E_{HS}$ and $E_{QM}$ are different.
+\emph{\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}}
 
-Guided by figure \ref{fig_map}, let us study the map $\iota$ by going through a few examples.
-Since every pure state in quantum mechanics has zero entropy, we will require the same for all pure states in $E_{HS}$.
-Let us consider the HS-epistemic states specified by three uniform distributions over the Bloch sphere with support given by:
-\begin{description}
-	\item[A] the poles, $z^+$ and $z^-$;
-	\item[B] the equator, all states on the x/y planes;
-	\item[C] the whole sphere, all points.
-\end{description}
-Since we want the entropy for a single pure state to be zero, the entropy for case \textbf{A} will be 1 (uniform distribution between two discrete cases).
-Since $\iota(A)$ will be the maximally mixed state, this also corresponds to entropy 1.
-So far so good: $\iota$ maps entropy 1 to entropy 1.
-However, note that \textbf{A} is a measure zero set of \textbf{B}, which is a measure zero set of \textbf{C}.
-The entropy of \textbf{C} is therefore infinite with respect to \textbf{B} which is infinite with respect to \textbf{A}.
-On the other hand, \textbf{A}, \textbf{B} and \textbf{C} will all correspond to the same maximally mixed state, therefore $\iota$ collapses together epistemic states at different levels of infinity.
-Note that, in this case, $\iota$ picks the minimum entropy between the equivalent epistemic states.
-This is inconsistent with the idea that $E_{QM}$ represents states where knowledge is lost: that would pick the maximum possible entropy.
+\emph{Proof}: Suppose we have a $\psi$-ontic model defined according to \cite{Harrigan:2010}.
+The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping distributions over a space $\Lambda$, therefore eq.
+\ref{entropy_nonoverlap} applies.
+Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states is zero, we must have
+\begin{equation}\label{entropy_pure}
+	H(p(\lambda|P_\psi)) = H(p(\lambda|P_\phi)) = 0,
+\end{equation}
+and therefore
+\begin{equation}\label{required_entropy}
+	\begin{aligned}
+		H\left(p(\lambda|\frac{1}{2}P_\psi + \frac{1}{2}P_\phi)\right) &= \\
+		H\left(\frac{1}{2}p(\lambda|P_\psi) + \frac{1}{2}p(\lambda|P_\phi)\right) 
+		&= 1.
+	\end{aligned}
+\end{equation}
+If we compare the above with eq.~\ref{entropy_mixed}, it follows that $p$ must be zero. That is, we must have that
+\begin{equation}\label{orthogonal}
+	\langle \psi | \phi \rangle = 0
+\end{equation}
+no matter what $\psi$ and $\phi$ are.
 
-\begin{figure}
-\includegraphics[scale=.4]{fig2}
-\caption{\footnotesize{The function $\iota$ maps the epistemic states $E_{HS}$ of a $\psi$-ontic model of a single qubit to the respective density matrices $E_{QM}$.
-Note that neither the value nor the ordering of the information entropy is preserved.}}\label{fig_map}
-\end{figure}
+Hence, the non-overlapping assumption built into the $\psi$-ontic model necessarily implies that all pure states are orthogonal.
+Since this is not true in quantum mechanics, any $\psi$-ontic model will fail to reproduce the results of quantum information, quantum statistical mechanics and, therefore, quantum theory in general.
+$\square$
 
-Now consider another uniform distribution over the support:
-\begin{description}
-	\item[D] the north pole and a point on the equator, $z^+$ and $x^+$.
-\end{description}
-This is yet another uniform distribution between two discrete cases, therefore in $E_{HS}$ the entropy is 1.
-However, $\iota(D)$ is not the maximally mixed state and therefore the entropy is less than 1.
-This means that epistemic states that are isoentropic in $E_{HS}$ are not isoentropic in $E_{QM}$.
-If we think of $E_{HS}$ and $E_{QM}$ as partially ordered by information entropy, then $\iota$ is not an order preserving map.
-Therefore, the epistemic structures represented by $E_{HS}$ and $E_{QM}$ are not isomorphic.
+We should also note that one has problems for $\psi$-epistemic models.
+While PBR leaves the possibility of $\psi$-epistemic models that use correlations, the correlations is also fixed by the entropy.
+It is a well-known result  [CITE: libri di informazione] in both classical and quantum information theory that the entropy of a joint distribution of the marginal if and only if the subsystems are independent.
+Therefore an epistemic state that represents independent quantum states must also be the joint distribution of statistically independent epistemic states of the individual systems.
+This, combined with PBR, tells us that no models can reproduce quantum statistical mechanics and quantum information theory.
 
-The seemingly innocuous assumption that $p(\lambda|P)$ is well defined for all preparations $P$ and for all states $\lambda$ imposes a classical probability space which follows the rules of classical information theory, and prevents the state space $\Lambda$ from supporting an epistemic structure compatible with quantum theory.
+The conclusion is the following:
+\begin{quote}
+	\textbf{no ontological model can reproduce quantum statistical mechanics and quantum information with standard information theoretic tools}.
+\end{quote}
+That is, even if one had an ontological model that mapped the result of single outcome correctly, it would still give the wrong answer when trying to calculate ensembles and their entropy.
 
-This suggests an obvious follow up: what if we start by assuming that the set of pure states $\Lambda$ supports the correct epistemic structure, the one associated with density matrices, and use the same definitions of $\psi$-ontic and $\psi$-epistemic?
-We can rephrase condition \ref{ontic_condition} directly in terms of density matrices as $\rho_1 \rho_2 = 0$ which, for pure states, becomes:
-\begin{equation}
-	|\psi_1\rangle\langle\psi_1|\psi_2\rangle\langle\psi_2| = 0.
-\end{equation}
-That is, two pure states are non-overlapping if and only if $\langle\psi_1|\psi_2\rangle=0$, i.e.
-if they are orthogonal.
-Alternatively, we could ask that the probability of being in the support of the second given the first be zero.
-For pure states, we get $\langle\psi_2|\psi_1\rangle\langle\psi_1|\psi_2\rangle = 0$, which leads to the same condition.
-Or we could define the overlap in terms of the entropy of the mixture and still find the same result.
+In principle, we do not exclude that someone may find a different construction to reproduce the correct entropies.
+There are, however, at least two objective difficulties that cannot be circumvented.
+First of all, information entropy is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state.
+Therefore entropy fixed equality.
+Because the mapping between epistemic states and density matrices is not injective,\footnote{Quantum mixture do not have a single decomposition in pure states.
+For example, the maximally mixed state of a qubit can be achieved with either an equal mixture of spin up and down or an equal mixture of spin left and right.} epistemic states must form equivalence classes and the strict concavity must be lost.
+That is, one can make mixtures of different ensemble without increasing entropy.
 
-However we look at it, we reach the following result:
-\begin{align}\label{conclusion}
-	\parbox{2.8in}{Non-orthogonal pure quantum states overlap and are therefore HS-epistemic.}
-\end{align}
-This should not be surprising: even in classical probability two distributions are non-overlapping if and only if they are orthogonal with respect to the inner product $\int \rho_1 \rho_2 d\lambda$.
+Secondly, as equation \ref{entropy_mixed} shows, the compute of the entropy requires the inner product.
+Therefore, whatever other structure one puts on top of epistemic states, essentially redefines the inner product.
+In other words, the idea that the inner product just represent transition probabilities, which, in our view, is the key idea the ontological model uses to explain ``what really happens'' does not work.
 
-Therefore, we have the following result: if we want a set of pure states that is able to support the epistemic structure of QM, they must be ``overlapping'' and thus we must have a $\psi$-epistemic model.
-This may seem in contradiction with the PBR theorem.
-Nonetheless, we believe that there is no contradiction, since there is no model in the HS classification that is compatible with quantum theory.
-In other words, the categorization itself is flawed.
-
-\section{Discussion}
-
-Let us now ask a crucial question: Why does assuming that $p(\lambda|P)$ is well defined lead to an epistemic structure incompatible with QM?
-
-It is widely understood that one feature of quantum theory is contextuality: a measurement cannot be conceived as revealing pre-existing values.
-We may be tempted to think that we have one context before and one after the measurement.
-This does not work because we have a related dual effect on preparations.
-In the same way that we cannot understand quantum measurements as selecting from a single joint probability distribution for all properties, we cannot understand quantum preparations as constructing a single joint probability distribution over all properties.
-To fully understand how this works, let us review how probability spaces are defined in both classical and quantum mechanics.
-
-A ``classical'' probability space is made of three objects: a sample space $\Omega$ representing all possible cases (e.g.\ all points in phase-space for a classical system); a $\sigma$-algebra $\Sigma_\Omega$ over $\Omega$  representing all statements of interest (e.g.\ ``the position is between 2 and 3 meters''); a measure $\mu : \Sigma_\Omega \to \mathbb{R}$ associating each statement with a probability.
-Note how we have a single probability distribution over the whole space.
-
-In QM we start with a similar structure: a set of states $\mathcal{H}$ representing all possible cases (i.e.\ the Hilbert space) and a $\sigma$-algebra $\Sigma_{\mathcal{H}}$ representing all statements of interest.
-This is where the similarity ends.
-To specify a mixed state---a ``quantum distribution''---we use a density matrix $\rho : \mathcal{H} \to \mathcal{H}$.
-To get a probability distribution, we need to specify an observable $O$.
-The observable $O$ identifies an orthogonal basis and we take the sub-algebra $\Sigma_O \subset \Sigma_{\mathcal{H}}$ defined on those states.
-We can then associate $\Sigma_O$ with a measure $\mu_O : \Sigma_O \to \mathbb{R}$, which is therefore not on the whole space $\Sigma_{\mathcal{H}}$.
-
-Mathematically, what we call context is a $\sigma$-algebra: a set of propositions upon which we assign probabilities.
-This fact explains why we cannot simply use classical rules in QM.
-Classical mechanics has only one context, so we can always mix and match statements however we please.
-In QM, propositions about different observables live in different $\sigma$-algebras---in different probability spaces---and therefore cannot be combined.
-
-One may think that we can get away by assuming that, at each time, there is a ``true'' context.
-And this seems to work in some cases.
-Given any density matrix $\rho$, there exists a (non-unique) privileged context: this is the one in which we can express $\rho = \sum p_i |\psi_i \rangle \langle \psi_i|$ as the mix of orthogonal pure states.
-If $\rho$ is itself a pure state, any basis that includes it will be a privileged context.
-This privileged context provides us the probability distribution over which the information entropy is calculated.
-After a measurement, the privileged context changes to one that includes the eigenstates of the chosen observable.
-However, this does not work in general.
-
-Consider two density matrices $\rho_1$ and $\rho_2$ with different eigenstates and their mixture $\rho = p_1 \rho_1 + p_2 \rho_2$.
-The context of the mixture will not be the same as the original contexts.
-It will, in fact, even depend on $p_1$, $p_2$.
-That is, when mixing preparations we are not simply combining two probability distributions over the same context, like in classical mechanics, but we are combining the contexts themselves to create a new one.
-Contextuality not only affects measurements, but, dually, preparations as well.
-And this affects how information entropy is calculated and the overall epistemic structure.
-
-To sum up, assuming $p(\lambda|P)$ exists means assuming there is a single probability space, a single $\sigma$-algebra, a single set of propositions to which all probabilities are assigned.
-This is exactly assuming non-contextuality.
-While this can be made to work for pure state-to-pure state transitions during measurements, mixing preparations leads to an epistemic structure that is fundamentally non-contextual, leading to inconsistencies with quantum information.
-To allow for contextuality we should assume that for each observable $O$ we have a probability $p_O(\lambda_O|P)$ of having prepared one of the eigenstates, such that $\sum_{\lambda_O} p_O(\lambda_O|P) = 1$.
-Although we have a single state space $\Lambda$, we do not have a single probability space over the whole space.
-The entropy of the preparation would be the lowest entropy across all contexts.
+Furthermore, we stress that it is not our place to solve these problems: we are not the proponent of this model.
+Our only aim is to raise enough doubt that the ontological model, as it currently stand, is highly problematic.
 
@@
 %\bibliographystyle{ieeetr}
-\bibliography{HarriganSpekkens}
+\bibliography{bibliography}
+\clearpage
+
+\section*{Appendix: calculation}
+\label{A}
+\textbf{Entropy of mixed non-overlapping distributions}.
+We want to show that, given two non-overlapping probability distributions $\rho_1$ and $\rho_2$, the entropy of $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ is given by
+\begin{equation}
+	H(\rho) = 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2). \tag{\ref{entropy_nonoverlap}}
+\end{equation}
+
+Let $U_1, U_2 \subset \Lambda$ be the respective supports of the distributions. Since the distributions are non-overlapping, we have $U_1 \cap U_2 = \emptyset$. We have
+\begin{align*}
+	H(\rho) &= - \int_\Lambda \rho \log \rho d\lambda \\
+	&= -\int_{U_1} \rho \log \rho d\lambda -\int_{U_2} \rho \log \rho d\lambda \\
+	&= -\int_{U_1} \frac{1}{2} \rho_1 \log \frac{1}{2} \rho_1 d\lambda -\int_{U_2} \frac{1}{2} \rho_2 \log \frac{1}{2} \rho_2 d\lambda \\
+	&= - \frac{1}{2} \int_{U_1} \rho_1 \log \frac{1}{2} d\lambda - \frac{1}{2} \int_{U_1} \rho_1 \log \rho_1 d\lambda \\
+	&- \frac{1}{2} \int_{U_2} \rho_2 \log \frac{1}{2} d\lambda - \frac{1}{2} \int_{U_2} \rho_2 \log \rho_2 d\lambda \\
+	&= - \frac{1}{2} \log \frac{1}{2} - \frac{1}{2} \log \frac{1}{2} + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2) \\
+	&= 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2). \\
+\end{align*}
+
+\textbf{Entropy of quantum mixed states}.
+We want to show that, given two states $\psi$ and $\phi$, the entropy of the mixed state $\rho = \frac{1}{2}|\psi\rangle\langle\psi| + \frac{1}{2}|\phi\rangle\langle\phi|$ is
+\begin{equation}\label{entropy}
+	H(\rho) = H\left(\frac{1+|\langle\psi|\phi\rangle|}{2}, \frac{1-|\langle\psi|\phi\rangle|}{2}\right).
+\end{equation}
+
+\begin{center}
+	\begin{tikzpicture}[scale = 1]
+		\draw (0,0) circle (2);
+		\node at (-2.3,0) {$-$};
+		\node at (2.3,0) {$+$};
+		\node at (2,1.4) {$\psi$};
+		\node at (2,-1.4) {$\phi$};
+		\draw (-2,0) -- (2,0);
+		\begin{scope}
+			\clip(0,0) circle (2);
+			\draw (0,0) -- (2,1.4);
+			\draw (0,0) -- (2,-1.4);
+			\draw (1.634,1.4) -- (1.634,-1.4);
+		\end{scope}
+		\fill (1.634,0) circle (0.05);
+		\node at (1.8,.2) {$\rho$};
+	\end{tikzpicture}
+\end{center}
+
+Note that $\psi$ and $\phi$ will identify a two-dimensional subspace which can be thought, without loss of generality, as a qubit and therefore can be represented by a Bloch sphere.
+The picture represents the intersection of the Bloch sphere with the plane identified by $\psi$ and $\phi$.
+As $\rho$ is an equal mixture of the two states, it will be represent by the midpoint between the two.
+Taking the line that goes through $\rho$ and the center of the sphere, we can see that $\rho$ can also be seen as the mixture of the states $+$ and $-$ which, since they represent equal and opposite directions, form a basis.
+To diagonalize $\rho$, then, means to express it in terms of $+$ and $-$.
+
+If $\theta_{\psi\phi}$ is the angle between $\psi$ and $\phi$, we have
+\begin{equation}
+	|\langle \psi | \phi \rangle |^2 = \cos^2 \frac{\theta_{\psi\phi}}{2}.
+\end{equation}
+The angle is divided by two because the angle on the Bloch sphere (i.e.
+in physical space) is double the angle in the Hilbert space.
+For example, for $z^+$ and $z^-$ the angle on the Bloch sphere would be $\pi$ and the inner product is zero (i.e.
+opposite directions in physical space correspond to orthogonal states).
+
+Now we express $\psi$ and $\phi$ in terms of $+$ and $-$, remembering that they form a basis.
+Given that $\rho$ is at the midpoint, the figure is vertically symmetric.
+The angle between $\psi$ and $+$, then, is half of $\theta_{\psi\phi}$.
+The inner product between $\psi$ and $+$ is
+\begin{equation}
+	\begin{aligned}
+		|\langle \psi | + \rangle |^2 &= \cos^2 \frac{\theta_{\psi +}}{2} \\
+		&= \cos^2 \frac{\theta_{\psi\phi}}{4}.
+	\end{aligned}
+\end{equation}
+Keeping in mind that we are composing vectors in the Hilbert space (and not in the geometry of the physical space) we have
+\begin{align*}
+	\left|\psi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>+\sin\frac{\theta_{\psi\phi}}{4}\left|-\right> \\
+	\left|\phi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>-\sin\frac{\theta_{\psi\phi}}{4}\left|-\right>.
+\end{align*}
+
+The density matrices corresponding to the pure states are
+\begin{align*}
+	\left|\psi\right>\left<\psi\right|&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right|\\
+	&+\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
+	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right| \\
+	\left|\phi\right>\left<\phi\right|&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right|\\
+	&-\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
+	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right|.
+\end{align*}
+
+We can now calculate the mixture
+\begin{align*}
+	\frac{1}{2}(|\psi\rangle\langle\psi| &+ |\phi\rangle\langle\phi|) \\
+	&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right| +\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right| \\
+	&=\frac{1+\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|+\right>\left<+\right| +\frac{1-\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|-\right>\left<-\right| \\
+	&=\frac{1+|\langle\psi|\phi\rangle|}{2}\left|+\right>\left<+\right| +\frac{1-|\langle\psi|\phi\rangle|}{2}\left|-\right>\left<-\right|. \\
+\end{align*}
+
+As $\rho$ is in a diagonal form, the entropy is given by eq.~\ref{entropy}.
 

### 2023-02-09 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
ClassMechPhil: version submitted to Erke

--- 2021-08-05 Articles/2020-ClassicalMechanicsPhil/Submissions/PoS/classical-phil-article-PoS.tex
@@
 %\geometry{top=4cm, bottom=4cm, left=3cm, right=4cm} %close to actual philsci margins
+
+%\usepackage{endnotes}
+%\let\footnote=\endnote %Erkenntnis requires end-notes, but recent papers don't seem to follow this in the pdf!
+ % %To call the endnotes, use command \theendnotes before calling reference page
 
@@
 \usepackage{titlesec}
+
+\titleformat*{\section}{\large\bfseries}
+\titleformat*{\subsection}{\normalsize\bfseries}
+\titleformat*{\subsubsection}{\normalsize\bfseries}
+
 \iffalse
 \titleformat{\section}[runin]% runin puts it in the same paragraph
-        {\normalfont\bfseries}% formatting commands to apply to the whole heading
+        {\normalsize\bfseries}% formatting commands to apply to the whole heading
         {\thesection.}% the label and number
@@
 
-\titlespacing*{\section}{0pt}{5mm}{2.5mm}%spacing from left margin, spacing above header, spacing on right of header
+\titlespacing*{\section}{0pt}{2.5mm}{2.5mm}%spacing from left margin, spacing above header, spacing on right of header
 \fi
@@
 \fancyhead[R]{\thepage} %page number placement
-\fancyhead[C]{\textsc{HAMILTONIAN PRIVILEGE}}% odd page header 
+\fancyhead[C]{\textsc{Hamiltonian Privilege}}
 %\fancyhead[EC]{\textsc{AUTHOR NAMES}}%Even page header 
@@
 
+@Article{Gilton2021,
+  Author    = {Gilton, Marian J. R.},
+  Title     = {Could Charge and Mass be Universals?},
+  Journal   = {Philosophy and Phenomenological Research},
+  Year      = {2021},
+  Volume    = {102},
+  Pages     = {624--644},
+  Publisher = {Wiley Online Library},
+Url = {https://doi.org/10.1111/phpr.12681}, 
+}
+
 \end{filecontents}
@@
 
-\DeclareFieldFormat{pages}{\mkcomprange{#1}} % compresses page range in bibliography, so that unchanged digits are not repeated.
+%\DeclareFieldFormat{pages}{\mkcomprange{#1}} % compresses page range in bibliography, so that unchanged digits are not repeated.
 
 \DefineBibliographyStrings{english}% for your language
-{ byeditor = {ed.}, } %changes `Edited by' to `ed.' See Stanley easier guide, p. 35
+{ byeditor = {(Ed.)}, } %changes `Edited by' to `ed.' See Stanley easier guide, p. 35
 
@@
 
-\newif\iffootpunct %defines \footpunctfalse  and \footpuncttrue commands
-
-% % Essential to modify footnote ONLY AFTER BIBLATEX COMMANDS, since otherwise ruins citation styles
-
-% % footnote modification: must be between \makeatletter and \makeatother
- \makeatletter
-
- \def\blfootnote{\gdef\@thefnmark{} \@footnotetext} %footnote without number marker
-
- \renewcommand{\@makefntext}[1]{%
-  \setlength{\parindent}{0pt}%
-  \begin{list}{}{\setlength{\labelwidth}{0mm}% indents footnote text, Space from number to border
-    \setlength{\leftmargin}{\labelwidth}%  
-    \setlength{\labelsep}{0pt}% Space from number to text, moves footnote mark to the left
-     \setlength{\itemsep}{0pt}%
-      \setlength{\parsep}{0pt}%
-      \setlength{\topsep}{0pt}% removes separation between footnotes
-      }%\doublespacing \normalsize% double spaces and changes font size of footnotes; place before }
-  \item[\@thefnmark \iffootpunct. \fi \hfil ]#1% @makefnmark, with period iff footpuncttrue
-  \end{list}%
-}
-
-%      \setlength{\footnotesep}{0.83125\baselineskip}% separation between footnotes
-
-\makeatother
-
-%\renewcommand{\footnotelayout}{\normalsize}
-
 \begin{document}
 
-\title{\vspace{-3em} \LARGE Hamiltonian Privilege }			
+\title{\vspace{-3em} \LARGE Hamiltonian Privilege}			
 %\author{\Large \textsuperscript{*\textdagger}}
-\date{\vspace{-7.9ex} July 30, 2021}
+\date{\vspace{-7.9ex} April 3rd, 2022}
 \maketitle
@@
 
-\noindent\rule{\textwidth}{1.25pt}
+%\noindent\rule{\textwidth}{1.25pt}
 \bs
+\begin{flushleft}{\large \textbf{Abstract}}\end{flushleft}
+
 \noindent {\normalsize We argue that Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
-Our argument provides a non-metaphysical strategy for privileging one formulation of a theory over another: a formulation is more fundamental when it is more general.
+Our argument provides a non-metaphysical strategy for privileging one formulation of a theory over another: \textit{ceteris paribus}, a more general formulation is more fundamental.
 We illustrate this criterion through a novel interpretation of classical mechanics, based on three physical conditions.
@@
 \es
-\noindent\rule{\textwidth}{1.25pt}
+%\noindent\rule{\textwidth}{1.25pt}
 
@@
 \es
-
-\footpunctfalse% turns off period after footnote marker for blindfootnote
-
-%\blfootnote{*To contact the author, please write to: }
-%\blfootnote{\textdagger For comments on earlier drafts, we thank }
-
-\footpuncttrue% turns back on the period after the footnotemarker
-
-\raggedright\setlength{\parindent}{1em}
 
@@
 
-In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must clarify what we mean by ``fundamental.'' Of various notions of fundamentality, we mean something entirely non-mysterious.
-Given two formalisms that capture a class of systems, if one is logically more general, then it is conceptually more fundamental.
-Conceptual fundamentality is familiar from mathematics.
+In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must provide a criterion for a framework being more fundamental than another.
+Of various notions of relative fundamentality, we focus on an entirely non-mysterious aspect.
+Given two formalisms that capture a class of systems, if one is logically more general, then it is more fundamental.
+This aspect of fundamentality is familiar from mathematics.
 A more fundamental framework encompasses a less fundamental one as a special case.
@@
 Similarly, an axiomatization which includes additional axioms is less fundamental than one which drops these axioms while maintaining the same expressive power.
-Conceptual fundamentality is distinct from notions of fundamentality associated with differences in energy scale, such as how the Standard Model of particle physics is more fundamental than classical mechanics.
+This aspect of fundamentality is distinct from aspects associated with differences in energy scale, such as how the Standard Model of particle physics is more fundamental than classical mechanics.
 We do not mean that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in this sense.
+As Gilton notes, ``many philosophers of science have moved away from thinking of fundamentality in science in terms of the very small toward thinking of it in terms of comprehensiveness, or breadth of applicability'' \parencites*[637]{Gilton2021}.
+%we must provide a criterion for one framework being more or less fundamental than another
 %\footnote{In Section~\ref{density}, we embrace a related non-metaphysical notion of fundamentality that pertains to physical entities rather than theoretical frameworks.}
 
-To apply our criterion for fundamentality, we must show that Lagrangian mechanics is less general than Hamiltonian mechanics.
+To apply our criterion for relative fundamentality, we must show that Lagrangian mechanics is less general than Hamiltonian mechanics.
 We do this in Section~\ref{derivation} by recovering the mathematical formalisms of both Hamiltonian and Lagrangian mechanics from three physical conditions.
-Our aim here is to show how this argument works in principle.
-Section~\ref{fundamentality} demonstrates some philosophical advantages of our preferred notion of fundamentality.\footnote{One might worry that our criterion for fundamentality is subject to elementary counterexamples.
+In this section, we describe how our argument works in principle.
+Section~\ref{fundamentality} demonstrates some philosophical advantages of our criterion for relative fundamentality.\footnote{One might worry that our criterion is subject to elementary counterexamples.
 For instance, can't we make any theory more general by adjoining arbitrary disjunctions?
@@
 First, we show that the formalism of Hamiltonian mechanics applies to systems that are both i) infinitesimally reducible and ii) follow a deterministic and reversible evolution.
-A system is \textit{infinitesimally reducible} provided that the state of the whole system is equivalent to specifying the states of its parts (and vice versa).
+A system is \textit{infinitesimally reducible} provided that specifying the state of the whole system is equivalent to specifying the states of its parts (and vice versa).
 Colloquially, we can say that the whole state equals the sum of its parts, where we take these infinitesimal parts to be classical particles.
@@
 We adopt our preferred terminology to avoid controversies surrounding how best to characterize separability.} We show in Section~\ref{infinitesimal} that the phase space of infinitesimally reducible systems must have at least the structure of a symplectic manifold.
-Recovering Hamiltonian mechanics then simply requires identifying when trajectories on phase space obey Hamiltonian evolution  (i.e.
+To recover Hamiltonian mechanics, it then suffices to identify when trajectories on phase space obey Hamiltonian evolution (i.e.
 evolve according to a symplectomorphism).
@@
 Physically, knowledge of the trajectory in physical space corresponds to knowledge of position and velocity, while knowledge of the trajectory in phase space requires knowing the system's momentum.
-Hence, motion equivalence entails a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics in the hyper-regular domain (Section~\ref{motion}).
+Hence, motion equivalence entails a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics (Section~\ref{motion}).
 As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
@@
 
-As indicated at the start of Section~\ref{privileging}, we use ``fundamentality'' in a non-metaphysical sense.
-Some differences in fundamentality are differences in generality: provided a theoretical framework still captures the relevant behavior, it is more fundamental if it requires fewer conditions to hold.
-Our notion of fundamentality leads to a correspondingly non-metaphysical notion of ``structure.'' Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
-Compared with North's \parencites*[]{North} account of structure, our definition has two advantages.
-First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which scientific anti-realists and pragmatists decry.
+As indicated at the start of Section~\ref{privileging}, we are interested in a non-metaphysical aspect of fundamentality.
+At least some differences in fundamentality are differences in generality: other things equal, a theoretical framework is more fundamental if it requires fewer conditions to hold.
+This aspect of fundamentality leads to a correspondingly non-metaphysical notion of structure.
+Hamiltonian mechanics employs less structure than Lagrangian mechanics because the former requires fewer physical conditions.
+Compared with North's \parencites*[]{North} account of comparative structure, our criterion has two advantages.
+First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which scientific anti-realists and pragmatists view with suspicion.
 Secondly, it avoids a technical objection that Swanson and Halvorson \parencites*[]{Swanson} levy against North's more precise mathematical criterion for comparing structure.
+%provided a theoretical framework captures the relevant behavior, it is more fundamental if it requires fewer conditions to hold.
 
@@
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).
-Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.\footnote{Weak equivalence corresponds to a general treatment of Lagrangian mechanics that does not presuppose Riemannian geometry.
+Hence, weak equivalence suffices for the existence of a hyper-regular Lagrangian.
+As discussed in Section~\ref{Curiel}, Lagrangian mechanics only yields a unique evolution in the hyper-regular domain.\footnote{Weak equivalence corresponds to a general treatment of Lagrangian mechanics that does not presuppose Riemannian geometry.
 As we discuss in Section~\ref{Curiel}, Equation~\ref{weak_equivalence} corresponds to the kinematical constraints central to Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics.}
+%; otherwise, the positions and velocities do not determine how the system evolves
 
@@
 
-\clearpage
+%\newpage
+
+%\theendnotes
+
+%\clearpage
+
+\bigskip
 
@@
 
-\defbibheading{myheading}[REFERENCES]{%
-\centering \normalfont\textsc{{#1}}%
+\defbibheading{myheading}[References]{%
+\centering \large \textbf{{#1}}%
 %\centering \footnotesize\textsc{{#1}}%

### 2023-02-10 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: review from Andrea

@@
-\documentclass[twocolumn,prl,floatfix,superscriptaddress]{revtex4-2}
-\usepackage{graphicx, subfigure, amsfonts, amssymb, amsmath, array, gensymb, fontenc, color, lipsum}
-%\documentclass[aps,twocolumn,showpacs,preprintnumbers]{revtex4}
-%\usepackage{graphicx}  % Include figure files
-%\usepackage{subfigure}
-%\usepackage{multirow}
-%\linespread{1.0}
-%\usepackage{fancyhdr}
-%\usepackage{longtable}
-%\usepackage{parskip}
-%\usepackage[T1]{fontenc}
-%\usepackage{dcolumn}
-%\usepackage{bm}        
-%\usepackage{amsfonts}  
-%\usepackage{amsmath}   
-%\usepackage{amssymb}
+\documentclass[10pt,twocolumn, nofootinbib]{revtex4-1}
+%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
+%\documentclass[10pt,twocolumn,letterpaper]{article}
+
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{graphicx}
+\usepackage{amsfonts}
+\usepackage{enumitem}
+\usepackage{graphicx}
 \usepackage{hyperref}
 \usepackage{tikz}
-\hypersetup{colorlinks= true, allcolors=blue}
-%\setcitestyle{aysep={}}
-%\setlength{\parindent}{11pt}
+
+\hypersetup{
+	colorlinks=true,
+	citecolor=blue,
+	urlcolor=blue,
+	linkcolor=blue
+}
+\urlstyle{same}
 \frenchspacing
@@
 \newcommand\partitle[1]{\textsc{#1}.}
+
 \begin{document}
 
-\title{On the reality of the quantum state once again}
+\title{On the reality of the quantum state once again: A no-go theorem for $\psi$-ontic models}
 \author{Gabriele Carcassi}
@@
 \begin{abstract}
-The Pusey, Barrett and Rudolph (PBR) theorem claims that quantum states cannot be taken to represent mere information about the system.
-This result is based on the ontological model framework proposed by Harrigan and Spekkens (HS).
-In this paper we show that the HS framework has a fundamental problem: the epistemic structure it implicitly assumes does not follow the one dictated by quantum mechanics.
-Namely, the map between the epistemic states of the model and quantum density matrices preserves neither the value nor ordering of the information entropy.
-Consequently the epistemic content of mixed states is not mapped in a meaningful way.
-The problem stems from the assumption that an epistemic state is characterized by a \emph{single} probability measure, which is essentially an assumption of non-contextuality.
-Given this fundamental issue, every result that uses the HS framework, including the PBR theorem, should be carefully reexamined.
+In 2010 Harrigan and Spekkens (HS) proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic.
+Building on the HS framework Pusey, Barrett and Rudolph (PBR) proved a theorem claiming that $\psi$-epistemic models contradict quantum mechanics, and thus quantum states cannot be taken to represent mere information about quantum systems.
+In this essay we propose a simple no-go theorem that explains why also $\psi$-ontic models cannot be expected to reproduce quantum mechanics.
+We will see that, using information theoretic considerations, the lack of overlap of epistemic states requires all states to be orthogonal, which openly contradicts quantum theory.
+Combining our result with the PBR argument, our conclusion is that both $\psi$-epistemic and $\psi$-ontic models contradict quantum mechanics.
 \end{abstract}
@@
 
-In 2012 Pusey, Barrett and Rudolph published a formal result in \emph{Nature Physics}, widely known as the PBR theorem, showing that ``if the quantum state merely represents information about the real physical state of a system, then experimental predictions are obtained that contradict those of quantum theory'' (\cite{PBR:2012}, p.\ 475).
-Alternatively stated, PBR argued that in every model reproducing the statistics and predictions of Quantum Mechanics (QM) the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi$-ontic.
+In 2010 Harrigan and Spekkens (HS) proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}.
+While their original aim was to clarify Einstein's view of Quantum Mechanics (QM), the HS framework has been widely employed in the literature not only to categorize different formulations of QM, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Ladyman:2021} for critical discussions).
+
+Referring to this, one of the most influential works based on HS classification is due to Pusey, Barrett and Rudolph who published a formal result in \emph{Nature Physics}---widely known as the PBR theorem---showing that ``if the quantum state merely represents information about the real physical state of a system, then experimental predictions are obtained that contradict those of quantum theory'' (\cite{PBR:2012}, p.\ 475).
+Alternatively stated, PBR argued that in every model reproducing the statistics and predictions of QM the quantum state $\psi$ must represent real physical properties of the system under consideration and not agents' knowledge---i.e.\ models must be $\psi$-ontic.
 Consequently, quantum theories cannot be $\psi$-epistemic.
 
-Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information.
+Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information.
 On the other, scholars recently showed that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2021, DeBrota:2019}).
@@
 In this paper we are not going to rebut the theorem itself.
-We instead aim to carefully reexamine one of its premises, i.e.\ Harrigan and Spekkens' (HS) classification between $\psi$-ontic and $\psi$-epistemic ontological models  \cite{Harrigan:2010} on top of which the PBR result is derived.
-We believe that a careful examination of this assumption forces us to reevaluate the significance of the PBR conclusion.
+We instead aim to carefully re-examine one of its premises, i.e.\ HS classification between $\psi$-ontic and $\psi$-epistemic ontological models on top of which the PBR result is derived.
+We believe that a careful analysis of this assumption forces us to re-evaluate the significance of this classification as well as the meaning of the PBR conclusion.
+In other words, we re-examine the definitions and features of such models in order to understand whether they constitute a sound basis from which to draw conclusions on the nature of $\psi$, given the influence that the HS framework has in quantum foundations.
+As the reader will see, our analysis suggests that they are not.
 
-HS proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic.
-While the original aim of this classification was to clarify Einstein's view of quantum theory, the HS framework has been widely employed in the literature not only to categorize different interpretations, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Lewis:2012, Ladyman:2021} for critical discussions).
-Given the influence that the HS framework and the PBR theorem have in quantum foundations, it is worthwhile to look again at the former with more scrutiny in order to understand whether it is a sound basis from which to draw conclusions on the nature of the quantum state.
+While the PBR theorem is a no-go result for $\psi$-epistemic models, in this essay we provide a no-go theorem for $\psi$-ontic models: using information theoretic considerations we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM since the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all states must be orthogonal, which openly contradicts quantum theory.
+Therefore, we conclude, $\psi$-ontic models should not be employed in order to classify quantum theories.
 
-In this paper we will show two things:
-\begin{enumerate}
-	\item the information entropy as calculated on HS epistemic states does not match the one provided by quantum mechanics, therefore the epistemic structures provided by QM and HS are not isomorphic;
-	\item if we rephrase the overlapping condition directly in terms of density matrices, we find that non-orthogonal pure states ``overlap'', which rules out $\psi$-ontic models.
-\end{enumerate}
-Combining these results with PBR, our conclusion is that the HS classification itself is fundamentally problematic.
-The main issue lies in assuming that the epistemic state identifies a \emph{single} probability distribution over the whole set of ontic states.
-We show that this, in essence, is an assumption of non-contextuality, which is also inherited by the PBR theorem, casting doubts on its validity.
+Combining our result with the PBR argument, we conclude that the HS classification itself is fundamentally empty, i.e.\ both $\psi$-epistemic and $\psi$-ontic models do contradict quantum mechanics.
+Consequently, HS framework should be employed neither to classify interpretations of the quantum formalism, nor to draw conclusion about the nature of $\psi$.
 
@@
 In a \textbf{$\psi$-ontic} model the wave function is a physical property of the ontic state, in the sense that given an ontic state $\lambda$ there is only one pure state preparation $P_\psi$ that could have prepared $\lambda$.
-As we can see in figure \ref{overlap}, this happens if the probability distributions do not overlap, i.e.\ if we have
+As we can see in figure \ref{overlap} (a), this happens if the probability distributions do not overlap, i.e.\ if we have
 \begin{equation}\label{ontic_condition}
@@
 To reproduce quantum theory, an ontological model has to reproduce not simply the probability of transitions during measurements but also all the results of quantum statistical mechanics and quantum information theory.
-Therefore it has to reproduce the correct values for entropy of mixed states.
-The natural way is by using standard information theoretic tools over the space of epistemic states which, as we will see, rules out $\psi$-ontic models.
-The degree of overlap between distributions, in fact, is not a choice, but is fixed by the value of entropy.
+Therefore, among other things, it has to obtain the correct values for entropy of mixed states.
+\textcolor{blue}{The natural way is by using standard information theoretic tools over the space of epistemic states which, as we will see, rules out $\psi$-ontic models.
+The degree of overlap between distributions, in fact, is not a choice, but is fixed by the value of entropy.} %@Gabriele: expand these points, i.e.
+explain more clearly the inference that you have in mind.
 
-Let us first review a crucial property of the information entropy, both in the classical setting (i.e.\ the Shannon/Gibbs entropy) and in the quantum setting (i.e.\ the von Neumann entropy).\footnote{We show the explicit calculations in the appendix.}
+Let us first review a crucial property of the information entropy, both in the classical setting (i.e.\ the Shannon/Gibbs entropy) and in the quantum setting (i.e.\ the von Neumann entropy).\footnote{We show the explicit calculations in the appendix.} %@ Gabriele: why is it important to introduce both classical and quantum property of information entropy?
+Questi sono tutti commenti anti-referee.
+Esplicita meglio a cosa servono queste cose.
 
@@
 
-\emph{\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}}
+\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}
 
 \emph{Proof}: Suppose we have a $\psi$-ontic model defined according to \cite{Harrigan:2010}.
-The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping distributions over a space $\Lambda$, therefore eq.
-\ref{entropy_nonoverlap} applies.
+The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping distributions over a space $\Lambda$,  \textcolor{blue}{therefore eq.
+\ref{entropy_nonoverlap} applies}.
+%@Gabriele: mi rendo conto che  stupido ma  un commento anti-referee.
+Tu dici che l'eq.
+7 che riguarda distribuzioni classiche si applica al nostro caso che riguarda stati quantistici.
+Esplicita a prova di referee perchŽ si applica l'eq.
+7 e non 8
 Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states is zero, we must have
@@
 While PBR leaves the possibility of $\psi$-epistemic models that use correlations, the correlations is also fixed by the entropy.
-It is a well-known result  [CITE: libri di informazione] in both classical and quantum information theory that the entropy of a joint distribution of the marginal if and only if the subsystems are independent.
+%Commenti anti-reviewer: what do you mean here exactly when you speak about correlations?
+Do you know the existence of one of these models?
+It is a well-known result in both classical and quantum information theory that \textcolor{red}{the entropy of a joint distribution of the marginal if and only if the subsystems are independent} (\cite{Ash:2010}, \cite{Nielsen:2010}).
+%@Gabriele: a verb seems to be missing in the red sentence.
 Therefore an epistemic state that represents independent quantum states must also be the joint distribution of statistically independent epistemic states of the individual systems.
@@
 
-The conclusion is the following:
+Thus, one concludes the following:
 \begin{quote}
@@
 First of all, information entropy is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state.
-Therefore entropy fixed equality.
+\textcolor{blue}{Therefore entropy fixed equality}.
+%spiega meglio
 Because the mapping between epistemic states and density matrices is not injective,\footnote{Quantum mixture do not have a single decomposition in pure states.
@@
 Therefore, whatever other structure one puts on top of epistemic states, essentially redefines the inner product.
-In other words, the idea that the inner product just represent transition probabilities, which, in our view, is the key idea the ontological model uses to explain ``what really happens'' does not work.
+In other words, the idea that the inner product just represents transition probabilities, which, in our view, is the key idea the ontological model uses to explain ``what really happens'' does not work.
 
-Furthermore, we stress that it is not our place to solve these problems: we are not the proponent of this model.
-Our only aim is to raise enough doubt that the ontological model, as it currently stand, is highly problematic.
+It is worth stressing, however, that it is not our task to solve these problems since we are neither the proponents of this model, nor supporters of it.
+Our only aim is to underline that the HS framework, as it currently stand, is highly problematic to draw any conclusion about the interpretation of quantum theory and the nature of $\psi$.
 
@@
 
-We showed not only that the HS framework does not correctly describe the epistemic structure of QM, but also that it implicitly assumes non-contextuality.
-Hence, it cannot be employed to draw conclusions on the nature of quantum states.
-Consequently, any other result relying on the HS categorization inherits the same issues.
-In particular, it follows that the PBR theorem is derived from a technically ill-defined basis; thus, it is demonstrated neither that quantum states must necessarily be HS-$\psi$-ontic, nor that epistemic interpretations are contradicted by quantum theory.
-Concluding, we believe that while work in quantum foundations typically focuses on measurements and their effect on pure states, equal importance must be given to preparations and their effect on mixed states.
-The different behavior of entropy in classical and quantum mechanics has profound implications not just for information theory, but also for statistical mechanics and its measurable predictions.
+In this paper we proved that $\psi$-ontic models are not compatible with quantum information theory, and therefore with quantum mechanics itself.
 
-%A full solution to the ``measurement problem'' will therefore have to address the dual ``preparation problem.''
+One may think that the issue could be circumvented by substituting the Shannon entropy formula. %with what? 
+That is, the problem would not be with the non-overlapping assumption made by HS but rather with how the entropy is calculated.
+However, the link between probability theory, information theory and measure theory is so tight that it makes this objection untenable.
+If we take a set of states $U$ with a measure $\mu(U)$ that quantifies the states in the set, the probability of each state for a uniform distribution is given by $\rho = 1 / \mu(U)$ and the entropy by $\log \mu(U)$.
+The measure used to count states, the probabilities assigned to states and the entropy assigned to distributions are not independent mathematical structures: failure of one entails the failure of all.
+We will look deeper into this issue in a later work.
 
-%Number of words for the conclusion: 164
+Finally, combining our result with the PBR theorem it follows that both $\psi$-ontic and $\psi$-epistemic models cannot reproduce QM, contradicting the results and the predictions of quantum theory.
+This fact shows in turn that HS categorization of ontological models is fundamentally empty.
+Thus, such a framework should be employed neither to classify quantum interpretations, nor to draw conclusions on the nature of the quantum state.
+
+Referring to this, being the HS definitions inadequate to reproduce QM, it follows that any result derived from them rests on a technically ill-defined basis, as for instance the PBR theorem itself.
+If this is true, then, it is not demonstrated (i) that quantum states must necessarily be HS-$\psi$-ontic, and (ii) that HS-$\psi$-epistemic interpretations are contradicted by quantum theory.
+Concluding, we believe that the PBR theorem does not make claims on the reality of the quantum state, but it rather shows that HS definition of $\psi$-epistemic model contradicts QM.
+If interpreted in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories.
+Argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.
 
@@
 
-Let $U_1, U_2 \subset \Lambda$ be the respective supports of the distributions. Since the distributions are non-overlapping, we have $U_1 \cap U_2 = \emptyset$. We have
+Let $U_1, U_2 \subset \Lambda$ be the respective supports of the distributions, which are non-overlapping i.e.\ $U_1 \cap U_2 = \emptyset$. We then have
 \begin{align*}
@@
 The picture represents the intersection of the Bloch sphere with the plane identified by $\psi$ and $\phi$.
-As $\rho$ is an equal mixture of the two states, it will be represent by the midpoint between the two.
+As $\rho$ is an equal mixture of the two states, it will be represented by the midpoint between the two.
 Taking the line that goes through $\rho$ and the center of the sphere, we can see that $\rho$ can also be seen as the mixture of the states $+$ and $-$ which, since they represent equal and opposite directions, form a basis.
@@
 
-As $\rho$ is in a diagonal form, the entropy is given by eq.~\ref{entropy}.
+As $\rho$ is in a diagonal form, the entropy is given by \ref{entropy}.
 

### 2023-02-10 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: another pass

@@
 
-In this paper we are not going to rebut the theorem itself.
-We instead aim to carefully re-examine one of its premises, i.e.\ HS classification between $\psi$-ontic and $\psi$-epistemic ontological models on top of which the PBR result is derived.
-We believe that a careful analysis of this assumption forces us to re-evaluate the significance of this classification as well as the meaning of the PBR conclusion.
-In other words, we re-examine the definitions and features of such models in order to understand whether they constitute a sound basis from which to draw conclusions on the nature of $\psi$, given the influence that the HS framework has in quantum foundations.
-As the reader will see, our analysis suggests that they are not.
+While the discussion has been focused mainly on whether $\psi$-epistemic models are problematic or whether the PBR argument itself is problematic, we ask a different question: what if the HS classification itself is problematic?
+In particular, the framework assumes that once a full description of the transition probability is given for pure states, mixed states are constructed simply as a classical statistical mixture.
+But classical mixtures and quantum mixtures have different properties.
+For example, classical mixtures have a single decomposition in terms of pure states while quantum mixtures do not.
+Studying the behavior of states under mixing, then, opens a new venue of possible violations that are not as easily seen by looking at transition probability.
 
-While the PBR theorem is a no-go result for $\psi$-epistemic models, in this essay we provide a no-go theorem for $\psi$-ontic models: using information theoretic considerations we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM since the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all states must be orthogonal, which openly contradicts quantum theory.
-Therefore, we conclude, $\psi$-ontic models should not be employed in order to classify quantum theories.
+The core insight is that the degree of overlap between two probability distribution affects how the entropy of their mixture relates to the entropy of the components.
+Consequently, we find that the lack of overlap required by $\psi$-ontic models forces all pure states to be orthogonal.
+Since this openly contradicts quantum theory, we conclude that $\psi$-ontic models are not viable.
 
-Combining our result with the PBR argument, we conclude that the HS classification itself is fundamentally empty, i.e.\ both $\psi$-epistemic and $\psi$-ontic models do contradict quantum mechanics.
-Consequently, HS framework should be employed neither to classify interpretations of the quantum formalism, nor to draw conclusion about the nature of $\psi$.
+Combining our result with the PBR argument, we conclude that the HS classification itself is fundamentally flawed: both $\psi$-epistemic and $\psi$-ontic models do contradict quantum mechanics.\footnote{In principle, PBR leaves open the possibility of $\psi$-epistemic models that violate statistical independence.
+Given that statistical independence is linked to entropy additivity, that loophole can be closed as well.} Consequently, HS framework should be employed neither to classify interpretations of the quantum formalism, nor to draw conclusion about the nature of $\psi$.
 
@@
 
-To reproduce quantum theory, an ontological model has to reproduce not simply the probability of transitions during measurements but also all the results of quantum statistical mechanics and quantum information theory.
-Therefore, among other things, it has to obtain the correct values for entropy of mixed states.
-\textcolor{blue}{The natural way is by using standard information theoretic tools over the space of epistemic states which, as we will see, rules out $\psi$-ontic models.
-The degree of overlap between distributions, in fact, is not a choice, but is fixed by the value of entropy.} %@Gabriele: expand these points, i.e.
-explain more clearly the inference that you have in mind.
+The exposition of the HS classification makes it seem one has complete freedom in choosing the overlap between the overlap between epistemic states.
+In this section we will show that this is not the case: this is constrained by entropy of mixtures.
+These constraints cannot be replicated by a $\psi$-ontic model with standard information theoretic techniques.
+Since quantum statistical mechanics and quantum information theory depend on the correct value of entropy on mixtures, $\psi$-ontic models will violate quantum theory.
 
-Let us first review a crucial property of the information entropy, both in the classical setting (i.e.\ the Shannon/Gibbs entropy) and in the quantum setting (i.e.\ the von Neumann entropy).\footnote{We show the explicit calculations in the appendix.} %@ Gabriele: why is it important to introduce both classical and quantum property of information entropy?
-Questi sono tutti commenti anti-referee.
-Esplicita meglio a cosa servono queste cose.
+To show the violation, we will first study how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy and the von Neumann entropy.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic state, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
+The second will be used to calculate the entropy on the quantum case.
 
-\textbf{Classical information entropy of mixed non-overlapping distributions}.
-Let $\rho_1$ and $\rho_2$ be two classical distributions over a space $\Lambda$ with measure $\lambda$.
+\textbf{Information entropy of mixed non-overlapping probability distributions}.
+Let $\rho_1$ and $\rho_2$ be two probability density function over a space $\Lambda$ with measure $\lambda$.
+That is, they are the Radon-Nikodym derivative of the respective probability measures and the measure $\lambda$.
 The entropy for each distribution is given by the usual formula\footnote{The logarithm is assumed to be in base 2.}, for example
@@
 \end{equation}
-Suppose the two distributions are disjoint, and let $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ be a uniform mixture of the two distributions. The entropy of $\rho$ is given by
+Suppose the two distributions are disjoint, and let $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ be a uniform mixture of the two distributions.
+More precisely, $\rho$ is the Radon-Nikodym derivative of the average of the two probability measures with respect to $\lambda$.
+The entropy of $\rho$ is given by
 \begin{equation}\label{entropy_nonoverlap}
@@
 \end{equation}
+where the right side of the equation is the Shannon entropy calculated on the given values.
+Therefore the entropy of an equal mixture of two pure states depends only on the probability of measuring one having prepared the other.
 
@@
 \emph{Proof}: Suppose we have a $\psi$-ontic model defined according to \cite{Harrigan:2010}.
-The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping distributions over a space $\Lambda$,  \textcolor{blue}{therefore eq.
-\ref{entropy_nonoverlap} applies}.
-%@Gabriele: mi rendo conto che  stupido ma  un commento anti-referee.
-Tu dici che l'eq.
-7 che riguarda distribuzioni classiche si applica al nostro caso che riguarda stati quantistici.
-Esplicita a prova di referee perchŽ si applica l'eq.
-7 e non 8
-Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states is zero, we must have
+The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping probability distributions over a space $\Lambda$.
+That is, they are the two probability density functions defined to be the Radon-Nikodym derivatives between the respective probability measures over $\Lambda$ and the measure $\lambda$.
+Equation \ref{epistemic_mixing} tells us that the probability measure combine in the usual way, and therefore an equal mixture of the two states must obey \ref{entropy_nonoverlap}.
+Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states in quantum mechanics is zero, we must have
 \begin{equation}\label{entropy_pure}
@@
 \end{equation}
-If we compare the above with eq.~\ref{entropy_mixed}, it follows that $p$ must be zero. That is, we must have that
+If we compare the above with eq.~\ref{entropy_mixed}, it follows that $p$ must be zero.
+In fact, the only case the Shannon entropy for a two element distribution is equal to one is for the uniform distribution $H(1/2, 1/2) = 1$.
+That is, we must have that
 \begin{equation}\label{orthogonal}
@@
 
-We should also note that one has problems for $\psi$-epistemic models.
-While PBR leaves the possibility of $\psi$-epistemic models that use correlations, the correlations is also fixed by the entropy.
-%Commenti anti-reviewer: what do you mean here exactly when you speak about correlations?
-Do you know the existence of one of these models?
-It is a well-known result in both classical and quantum information theory that \textcolor{red}{the entropy of a joint distribution of the marginal if and only if the subsystems are independent} (\cite{Ash:2010}, \cite{Nielsen:2010}).
-%@Gabriele: a verb seems to be missing in the red sentence.
+If we combine this result with PBR, the only models that are not ruled our are $\psi$-epistemic models where the objects are not independently prepared.
+That is, where the joint distribution over ontic states is not the product of the marginals.
+However, this loophole can also been closed with information theoretic considerations.
+It is a well-known result in both classical and quantum information theory that the entropy of a joint distribution is the sum of the marginal if and only if the subsystems are independent (\cite{Ash:2010}, \cite{Nielsen:2010}).
 Therefore an epistemic state that represents independent quantum states must also be the joint distribution of statistically independent epistemic states of the individual systems.
-This, combined with PBR, tells us that no models can reproduce quantum statistical mechanics and quantum information theory.
+This tells us that $\psi$-epistemic models must represent composite states of independently prepared systems with independently distributed epistemic states, and therefore PBR applies.
+This closes the last loophole: no model in the HS categorization can reproduce quantum theory in its entirety, which includes quantum information and quantum statistical mechanics.
 
@@
 
+\section{Throwing out information theory}
+
+Note that our conclusion relies on the use of standard information theoretic tools.
 In principle, we do not exclude that someone may find a different construction to reproduce the correct entropies.
-There are, however, at least two objective difficulties that cannot be circumvented.
+That is, one may give a different definition of entropy to use on the ontic space.
+We believe this to be highly unlikely as there are, at least, two objective difficulties that cannot be circumvented.
+
 First of all, information entropy is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state.
-\textcolor{blue}{Therefore entropy fixed equality}.
+That is, we can check whether two states $\rho_1$ and $\rho_2$ are the same state simply by creating a mixture $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ and see whether we end up with the same entropy or not.
+In other words: \emph{entropy fixes equality}.
 %spiega meglio
-Because the mapping between epistemic states and density matrices is not injective,\footnote{Quantum mixture do not have a single decomposition in pure states.
-For example, the maximally mixed state of a qubit can be achieved with either an equal mixture of spin up and down or an equal mixture of spin left and right.} epistemic states must form equivalence classes and the strict concavity must be lost.
-That is, one can make mixtures of different ensemble without increasing entropy.
+The issue here is that the mapping between epistemic states and density matrices is not injective.
+Quantum mixtures, in fact, do not have a single decomposition in pure states.
+For example, the maximally mixed state of a qubit can be achieved with either an equal mixture of spin up and down or an equal mixture of spin left and right.
+Therefore the same quantum mixture must correspond to an equivalence class of epistemic states, which means whatever definition we have on entropy for epistemic states must not satisfy strict concavity to be able to agree with the quantum case.\footnote{The degree of degeneracy one has to require is also extreme.
+For a qubit, in the simplest case, we would have a $\psi$-ontic state for each direction of spin, meaning the space of ontic states is the sphere $S^2$.
+The space of all possible epistemic states is the set of all possible distribution integrable to one.
+This is an infinite dimensional function space.
+However, the space of mixed state for a quibit is the interior of a sphere, which is a bounded three-dimensional manifold.
+The equivalence classes of epistic states, then, must be infinite dimensional.} That is, one can make mixtures of different ensemble without increasing entropy, which does not make physical sense.
 
-Secondly, as equation \ref{entropy_mixed} shows, the compute of the entropy requires the inner product.
+Secondly, as equation \ref{entropy_mixed} shows, computing the entropy requires the inner product.
+In fact, if we had the entropy for all even mixtures of all pairs of pure states, we could reconstruct the inner product.
 Therefore, whatever other structure one puts on top of epistemic states, essentially redefines the inner product.
+Also, it has to do it in a way compatible with the other definitions of the model, therefore this structure will likely be ad-hoc for each model.
 In other words, the idea that the inner product just represents transition probabilities, which, in our view, is the key idea the ontological model uses to explain ``what really happens'' does not work.
 
-It is worth stressing, however, that it is not our task to solve these problems since we are neither the proponents of this model, nor supporters of it.
-Our only aim is to underline that the HS framework, as it currently stand, is highly problematic to draw any conclusion about the interpretation of quantum theory and the nature of $\psi$.
+One, naturally, will object that we haven't strictly proven that any of this is impossible or that there is no reasonable way to do it.
+Granted.
+However, it is not up to us to show that HS models cannot be made to work outside of the scope of standard information theory.
+And in any case, we believe we put forth enough evidence that the HS framework has serious flaws which, at the very least, must be understood by those that use it to draw any conclusion about the interpretation of quantum theory and the nature of $\psi$.
 
@@
 
-In this paper we proved that $\psi$-ontic models are not compatible with quantum information theory, and therefore with quantum mechanics itself.
-
-One may think that the issue could be circumvented by substituting the Shannon entropy formula. %with what? 
-That is, the problem would not be with the non-overlapping assumption made by HS but rather with how the entropy is calculated.
-However, the link between probability theory, information theory and measure theory is so tight that it makes this objection untenable.
-If we take a set of states $U$ with a measure $\mu(U)$ that quantifies the states in the set, the probability of each state for a uniform distribution is given by $\rho = 1 / \mu(U)$ and the entropy by $\log \mu(U)$.
-The measure used to count states, the probabilities assigned to states and the entropy assigned to distributions are not independent mathematical structures: failure of one entails the failure of all.
-We will look deeper into this issue in a later work.
-
-Finally, combining our result with the PBR theorem it follows that both $\psi$-ontic and $\psi$-epistemic models cannot reproduce QM, contradicting the results and the predictions of quantum theory.
-This fact shows in turn that HS categorization of ontological models is fundamentally empty.
+By using standard information theoretic techniques, we have shown that $\psi$-ontic models are not compatible with quantum information theory, and therefore with quantum mechanics itself.
+Given that the PBR theorem already ruled out $\psi$-epistemic models, the HS categorization of ontological models is fundamentally empty.
 Thus, such a framework should be employed neither to classify quantum interpretations, nor to draw conclusions on the nature of the quantum state.
 
-Referring to this, being the HS definitions inadequate to reproduce QM, it follows that any result derived from them rests on a technically ill-defined basis, as for instance the PBR theorem itself.
-If this is true, then, it is not demonstrated (i) that quantum states must necessarily be HS-$\psi$-ontic, and (ii) that HS-$\psi$-epistemic interpretations are contradicted by quantum theory.
-Concluding, we believe that the PBR theorem does not make claims on the reality of the quantum state, but it rather shows that HS definition of $\psi$-epistemic model contradicts QM.
-If interpreted in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories.
+The key problem of the HS categorization is that it is ``not quantum enough''.
+The framework tries to hide the non-classicality at the level of each ontic state, which, by construction, cannot be decomposed further using a standard probability measure.
+However, it posits that statistical mixtures are simply classical mixture in the sense that they are captured by a standard probability measure.
+But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.
+Therefore the HS categorization prevents one to describe quantum statistical mixing.
+
+Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are not sustainable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions.
+If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories.
 Argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.

### 2023-02-11 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: reviewed with Christine

@@
 \begin{abstract}
-In 2010 Harrigan and Spekkens (HS) proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic.
-Building on the HS framework Pusey, Barrett and Rudolph (PBR) proved a theorem claiming that $\psi$-epistemic models contradict quantum mechanics, and thus quantum states cannot be taken to represent mere information about quantum systems.
-In this essay we propose a simple no-go theorem that explains why also $\psi$-ontic models cannot be expected to reproduce quantum mechanics.
-We will see that, using information theoretic considerations, the lack of overlap of epistemic states requires all states to be orthogonal, which openly contradicts quantum theory.
-Combining our result with the PBR argument, our conclusion is that both $\psi$-epistemic and $\psi$-ontic models contradict quantum mechanics.
+	In this paper we show that $\psi$-ontic models, as defined by Harrigan and Spekkens (HS), cannot reproduce quantum theory.
+Instead of focusing on probability, we use information theoretic considerations to show that all pure states of $\psi$-ontic models must be orthogonal to each other, in clear violation of quantum mechanics.
+Given that  (1) Pusey, Barrett and Rudolph (PBR) previously showed that $\psi$-epistemic models, as defined by HS, also contradict quantum mechanics, and (2) the HS categorization is exhausted by these two types of models, we conclude that the HS categorization itself is problematic as it leaves no space for models that can reproduce quantum theory.
 \end{abstract}
@@
 
-While the discussion has been focused mainly on whether $\psi$-epistemic models are problematic or whether the PBR argument itself is problematic, we ask a different question: what if the HS classification itself is problematic?
-In particular, the framework assumes that once a full description of the transition probability is given for pure states, mixed states are constructed simply as a classical statistical mixture.
+While the discussion has been focused mainly on whether $\psi$-epistemic models are problematic or whether the PBR argument itself is problematic, we ask a different question: what if the underlying HS classification itself is problematic?
+In particular, the framework assumes that once pure states are fully characterized, mixed states are constructed simply as a standard (i.e.
+classical) statistical mixture.
 But classical mixtures and quantum mixtures have different properties.
 For example, classical mixtures have a single decomposition in terms of pure states while quantum mixtures do not.
-Studying the behavior of states under mixing, then, opens a new venue of possible violations that are not as easily seen by looking at transition probability.
+Studying the behavior of states under mixing, then, provides a new way to explore possible violations.
 
-The core insight is that the degree of overlap between two probability distribution affects how the entropy of their mixture relates to the entropy of the components.
+Using this new approach, in this paper we are going to provide a no-go theorem for $\psi$-ontic models.
+The core insight is that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.
 Consequently, we find that the lack of overlap required by $\psi$-ontic models forces all pure states to be orthogonal.
-Since this openly contradicts quantum theory, we conclude that $\psi$-ontic models are not viable.
+Since this directly contradicts quantum theory, we conclude that $\psi$-ontic models are not viable.
 
-Combining our result with the PBR argument, we conclude that the HS classification itself is fundamentally flawed: both $\psi$-epistemic and $\psi$-ontic models do contradict quantum mechanics.\footnote{In principle, PBR leaves open the possibility of $\psi$-epistemic models that violate statistical independence.
-Given that statistical independence is linked to entropy additivity, that loophole can be closed as well.} Consequently, HS framework should be employed neither to classify interpretations of the quantum formalism, nor to draw conclusion about the nature of $\psi$.
+Combining our result with the PBR argument, we conclude that the HS classification itself is fundamentally flawed: both $\psi$-epistemic and $\psi$-ontic models contradict quantum mechanics.\footnote{In principle, PBR leaves open the possibility of $\psi$-epistemic models that violate statistical independence.
+Given that statistical independence is linked to entropy additivity, that loophole can be closed as well.} Consequently, the HS framework should be employed neither to classify interpretations of the quantum formalism, nor to draw conclusions about the nature of $\psi$.
 
@@
 
-The exposition of the HS classification makes it seem one has complete freedom in choosing the overlap between the overlap between epistemic states.
+The exposition of the HS classification makes it seem one has complete freedom in choosing the overlap between epistemic states.
 In this section we will show that this is not the case: this is constrained by entropy of mixtures.
@@
 
-To show the violation, we will first study how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy and the von Neumann entropy.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic state, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
+To show the violation, we will first study how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy and the von Neumann entropy.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
 The second will be used to calculate the entropy on the quantum case.
@@
 \textbf{Information entropy of mixed non-overlapping probability distributions}.
-Let $\rho_1$ and $\rho_2$ be two probability density function over a space $\Lambda$ with measure $\lambda$.
-That is, they are the Radon-Nikodym derivative of the respective probability measures and the measure $\lambda$.
+Let $\rho_1$ and $\rho_2$ be two probability density functions over a space $\Lambda$ with measure $\lambda$.\footnote{The notation of HS uses $\lambda$ as both the value and measure for integration, as often done in single variable integrals.} That is, they are the Radon-Nikodym derivatives of the respective probability measures with respect to the measure $\lambda$.
 The entropy for each distribution is given by the usual formula\footnote{The logarithm is assumed to be in base 2.}, for example
@@
 \begin{equation}\label{entropy_mixed}
-	H(\rho) = H\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right).
+	\begin{aligned}
+	H(\rho) &= H\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right) \\
+	&= - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} - \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2} ,
+	\end{aligned}
 \end{equation}
@@
 The epistemic states $p(\lambda|P_\psi)$ and $p(\lambda|P_\phi)$ consist of non-overlapping probability distributions over a space $\Lambda$.
-That is, they are the two probability density functions defined to be the Radon-Nikodym derivatives between the respective probability measures over $\Lambda$ and the measure $\lambda$.
-Equation \ref{epistemic_mixing} tells us that the probability measure combine in the usual way, and therefore an equal mixture of the two states must obey \ref{entropy_nonoverlap}.
+That is, they are two non-overlapping probability density functions defined to be the Radon-Nikodym derivatives between the respective probability measures over $\Lambda$ and the measure $\lambda$.
+Equation \ref{epistemic_mixing} tells us that the probability measures combine in the usual way, and therefore an equal mixture of the two states must obey equation \ref{entropy_nonoverlap}.
 Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states in quantum mechanics is zero, we must have
@@
 If we compare the above with eq.~\ref{entropy_mixed}, it follows that $p$ must be zero.
-In fact, the only case the Shannon entropy for a two element distribution is equal to one is for the uniform distribution $H(1/2, 1/2) = 1$.
+In fact, the only case in which the Shannon entropy for a two-element distribution is equal to one is for the uniform distribution $H(1/2, 1/2) = 1$.
 That is, we must have that
@@
 
-If we combine this result with PBR, the only models that are not ruled our are $\psi$-epistemic models where the objects are not independently prepared.
+If we combine this result with PBR, the only models that are not ruled out are $\psi$-epistemic models where the objects are not independently prepared.
 That is, where the joint distribution over ontic states is not the product of the marginals.
-However, this loophole can also been closed with information theoretic considerations.
+However, this loophole can also be closed with information theoretic considerations.
 It is a well-known result in both classical and quantum information theory that the entropy of a joint distribution is the sum of the marginal if and only if the subsystems are independent (\cite{Ash:2010}, \cite{Nielsen:2010}).
@@
 \begin{quote}
-	\textbf{no ontological model can reproduce quantum statistical mechanics and quantum information with standard information theoretic tools}.
+	\textbf{No ontological model can reproduce quantum theory}.
 \end{quote}
-That is, even if one had an ontological model that mapped the result of single outcome correctly, it would still give the wrong answer when trying to calculate ensembles and their entropy.
+Note that HS do not claim that all interpretations must adhere to their framework, stating (\cite{Harrigan:2010}, p.
+134), ``Therefore, to categorize any given interpretation of the quantum formalism as $\psi$-ontic or $\psi$-epistemic, it is first necessary to cast it into the mold of an ontological model.
+If an interpretation resists being so cast, then it cannot be fit into our categorization.'' The result is that a successful interpretation of quantum mechanics must resists being so cast.
 
@@
 Quantum mixtures, in fact, do not have a single decomposition in pure states.
-For example, the maximally mixed state of a qubit can be achieved with either an equal mixture of spin up and down or an equal mixture of spin left and right.
-Therefore the same quantum mixture must correspond to an equivalence class of epistemic states, which means whatever definition we have on entropy for epistemic states must not satisfy strict concavity to be able to agree with the quantum case.\footnote{The degree of degeneracy one has to require is also extreme.
-For a qubit, in the simplest case, we would have a $\psi$-ontic state for each direction of spin, meaning the space of ontic states is the sphere $S^2$.
-The space of all possible epistemic states is the set of all possible distribution integrable to one.
+For example, the maximally mixed state of a qubit for a spin-$1/2$ system can be achieved with either an equal mixture of spin up and down or an equal mixture of spin left and right.
+Therefore epistemic states cannot be mapped one-to-one to quantum density operators, but rather each density operator corresponds to an equivalence class of epistemic states.\footnote{The degree of degeneracy one has to require is also extreme.
+For a qubit, in the simplest case, we would have a $\psi$-ontic state for each direction on the Bloch sphere, meaning the space of ontic states is the sphere $S^2$.
+The space of all possible epistemic states is the set of all possible distributions integrable to one.
 This is an infinite dimensional function space.
-However, the space of mixed state for a quibit is the interior of a sphere, which is a bounded three-dimensional manifold.
-The equivalence classes of epistic states, then, must be infinite dimensional.} That is, one can make mixtures of different ensemble without increasing entropy, which does not make physical sense.
+However, the space of mixed states for a quibit is the interior of the Bloch sphere, which is a bounded three-dimensional manifold.
+The equivalence classes of epistemic states, then, must be infinite dimensional.} This degeneracy means that the notion of equality among epistemic states is different than the notion of equality among density operators.
+In turn, this means that whatever definition we have for entropy on epistemic states must not satisfy strict concavity, if we want the value of entropy to agree between epistemic states and density operators.
+\emph{Entropy cannot fix equality for epistemic states}.
+In other words, one can make mixtures of different epistemic states without increasing entropy, which is questionable on physics grounds.
 
 Secondly, as equation \ref{entropy_mixed} shows, computing the entropy requires the inner product.
-In fact, if we had the entropy for all even mixtures of all pairs of pure states, we could reconstruct the inner product.
-Therefore, whatever other structure one puts on top of epistemic states, essentially redefines the inner product.
+In fact, if we had the entropy for all uniform mixtures of all pairs of pure states, we could reconstruct the inner product.
+Therefore, whatever other structure one puts on top of epistemic states essentially redefines the inner product.
 Also, it has to do it in a way compatible with the other definitions of the model, therefore this structure will likely be ad-hoc for each model.
@@
 
-The key problem of the HS categorization is that it is ``not quantum enough''.
-The framework tries to hide the non-classicality at the level of each ontic state, which, by construction, cannot be decomposed further using a standard probability measure.
-However, it posits that statistical mixtures are simply classical mixture in the sense that they are captured by a standard probability measure.
+The key problem of the HS categorization is that it is ``not quantum enough.'' The framework tries to hide the non-classicality at the level of each ontic state, which, by construction, cannot be decomposed further using a standard probability measure.
+However, it posits that statistical mixtures are simply classical mixtures in the sense that they are captured by a standard probability measure.
 But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.
-Therefore the HS categorization prevents one to describe quantum statistical mixing.
+Therefore the HS categorization prevents one from describing quantum statistical mixing.
 
 Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are not sustainable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions.
-If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories.
-Argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.
+If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories, an argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.
 

### 2023-02-13 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: adding failure composite systems

@@
 Combining our result with the PBR argument, we conclude that the HS classification itself is fundamentally flawed: both $\psi$-epistemic and $\psi$-ontic models contradict quantum mechanics.\footnote{In principle, PBR leaves open the possibility of $\psi$-epistemic models that violate statistical independence.
-Given that statistical independence is linked to entropy additivity, that loophole can be closed as well.} Consequently, the HS framework should be employed neither to classify interpretations of the quantum formalism, nor to draw conclusions about the nature of $\psi$.
+Given that statistical independence is linked to entropy additivity, we will see that that loophole can be closed as well.} Consequently, the HS framework should be employed neither to classify interpretations of the quantum formalism, nor to draw conclusions about the nature of $\psi$.
 
@@
 
-If we combine this result with PBR, the only models that are not ruled out are $\psi$-epistemic models where the objects are not independently prepared.
-That is, where the joint distribution over ontic states is not the product of the marginals.
-However, this loophole can also be closed with information theoretic considerations.
-It is a well-known result in both classical and quantum information theory that the entropy of a joint distribution is the sum of the marginal if and only if the subsystems are independent (\cite{Ash:2010}, \cite{Nielsen:2010}).
-Therefore an epistemic state that represents independent quantum states must also be the joint distribution of statistically independent epistemic states of the individual systems.
-This tells us that $\psi$-epistemic models must represent composite states of independently prepared systems with independently distributed epistemic states, and therefore PBR applies.
-This closes the last loophole: no model in the HS categorization can reproduce quantum theory in its entirety, which includes quantum information and quantum statistical mechanics.
-
-Thus, one concludes the following:
-\begin{quote}
-	\textbf{No ontological model can reproduce quantum theory}.
-\end{quote}
-Note that HS do not claim that all interpretations must adhere to their framework, stating (\cite{Harrigan:2010}, p.
-134), ``Therefore, to categorize any given interpretation of the quantum formalism as $\psi$-ontic or $\psi$-epistemic, it is first necessary to cast it into the mold of an ontological model.
-If an interpretation resists being so cast, then it cannot be fit into our categorization.'' The result is that a successful interpretation of quantum mechanics must resists being so cast.
-
 \section{Throwing out information theory}
@@
 Note that our conclusion relies on the use of standard information theoretic tools.
-In principle, we do not exclude that someone may find a different construction to reproduce the correct entropies.
+In principle, this does not exclude that someone may find a different construction to reproduce the correct entropies.
 That is, one may give a different definition of entropy to use on the ontic space.
@@
 
-One, naturally, will object that we haven't strictly proven that any of this is impossible or that there is no reasonable way to do it.
-Granted.
-However, it is not up to us to show that HS models cannot be made to work outside of the scope of standard information theory.
-And in any case, we believe we put forth enough evidence that the HS framework has serious flaws which, at the very least, must be understood by those that use it to draw any conclusion about the interpretation of quantum theory and the nature of $\psi$.
+In other words, the use of standard probability measures to describe the mixing is problematic in the same way that the use of standard probability measures would be a problem to describe the probability of transition.
+If we step back, we can understand our result and the PBR result in a broader context.
+
+\section{The problem with ontological models}
+
+If we combine our result with PBR, the only models that are not ruled out are $\psi$-epistemic models where the objects are not independently prepared.
+That is, where the joint distribution over ontic states is not the product of the marginals.
+However, this loophole can also be closed with information theoretic considerations.
+It is a well-known result in both classical and quantum information theory that the entropy of a joint distribution is the sum of the marginal if and only if the subsystems are independent (\cite{Ash:2010}, \cite{Nielsen:2010}).
+Therefore an epistemic state that represents independent quantum states must also be the joint distribution of statistically independent epistemic states of the individual systems.
+This tells us that $\psi$-epistemic models must represent composite states of independently prepared systems with independently distributed epistemic states, and therefore PBR applies.
+This closes the last loophole: no model in the HS categorization can reproduce quantum theory in its entirety, which includes quantum information and quantum statistical mechanics.
+
+Thus, one concludes the following:
+\begin{quote}
+	\textbf{No ontological model can reproduce quantum theory}.
+\end{quote}
+Now, the result may seem, at first, preposterous. But once we analyze the mathematics in detail, we see that both results, ours and PBR, are fundamentally finding something, in retrospect, obvious.
+
+It is well established that classical probability theory cannot recover the probability of quantum transitions.
+So, how does the HS model try to accommodate this?
+Given that $p(\lambda|P)$ combines using classical probability theory, all the non-classicality must be in $p(k|\lambda, M)$.
+In fact, note that this second expression does not allow us to write the joint probability distribution for multiple observers; therefore, each ontic state is not partitioned further.
+Therefore if $\lambda$ is the state of a single quantum system, we can hide the inner product inside $p(k|\lambda, M)$.
+This is enough to hide our non-classical probability for pure states of single quantum systems, but not much more.
+
+In our result, we saw that the use of standard measure theory when creating mixtures creates a problem because the convex space of classical distributions is significantly different from the convex space of quantum mixed states.
+For this case, ontological models are ``not quantum enough.''
+
+In PBR, the issue is breaking a composite system into parts. They find  that if the probability measure on $\lambda$ is factorized, that is (\cite{PBR:2012}, eq. 5) 
+\begin{equation}
+	\int_{\Lambda} \dots \int_{\Lambda} p(k | \lambda_1, \cdots, \lambda_n) \mu_{x_1}(\lambda_1)\cdots\mu_{x_n}(\lambda_n) d\lambda_1 \cdots d\lambda_n ,
+\end{equation}
+there is a problem. The issue here is not the factorization. The issue here is breaking the composite system into parts, each with its own ontic state as in 
+\begin{equation}
+	\begin{aligned}
+	\int_\Lambda &p(k | \lambda) d\lambda = \\
+	&\int_{\Lambda_1} \dots \int_{\Lambda_n} p(k | \lambda_1, \cdots, \lambda_n) \mu(\lambda_1, \cdots, \lambda_n) d\lambda_1 \cdots d\lambda_n.
+	\end{aligned}
+\end{equation}
+This assumes that a composite system is describable using a standard classical probability measure, which, in turn, allows us to express the joint probability distribution of observables of different subsystems.
+This, in general, will fail because it is ``not quantum enough.''
+
+Therefore there are two different ways ontological models use classical measure theory: one in mixtures, highlighted by us; one in composite systems, highlighted by PBR.
+The use of classical measure theory is what drives the failure of the model.
+
+The moral of the story is
+\begin{quote}
+	\textbf{In the context of quantum mechanics, standard probability measures are allowed \emph{only} over measurement outcomes/orthogonal states.}
+\end{quote}
+This, in fact, is exactly what we found in this paper using information theoretic considerations. In retrospect, as we said, it should not be at all surprising.
+
+Note that HS do not claim that all interpretations must adhere to their framework, stating (\cite{Harrigan:2010}, p.
+134), ``Therefore, to categorize any given interpretation of the quantum formalism as $\psi$-ontic or $\psi$-epistemic, it is first necessary to cast it into the mold of an ontological model.
+If an interpretation resists being so cast, then it cannot be fit into our categorization.'' The result is that a successful interpretation of quantum mechanics must resist being so cast.
 
@@
 Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are not sustainable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions.
+Specifically, the use of standard probability theory in composite systems.
 If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories, an argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.

### 2023-02-13 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
ClassicalMechanicsPhil: adding comments for referee

@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.}
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.} The general idea is that this assumption tells us that the state of a system is captured by a statistical distribution over the state space of the parts.
+If the state space of the parts is a manifold, this must be differentiable so that it can support a density at each point.
+Moreover, given that this density must be a function of the state, and not of the coordinates, we find that the manifold must be symplectic.
 % % % For further discussion, see \textcites{shpmp}. [omit this for review].
@@
 
-When the system is characterized by continuous quantities, $\mathcal{S}$ is a manifold.
-For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
+When the system is characterized by continuous quantities, $\mathcal{S}$ can be taken to be a manifold.\footnote{The use of manifolds is, technically, a choice of how the physics is represented.
+One could make another choice, for example, by representing the state variables not as local coordinates but as an algebra of operators that act on the state.
+In this case, we believe one should be able to run a similar argument and find the analogue of a symplectic structure for the different mathematical representation.} For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
 \textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called ``coordinates'' and $\xi$ is a coordinate map.
@@
 \noindent In our case, this applies to both the state of the whole system and the state of its parts.
+Under this assumption we will find the symplectic structure must be preserved in time, which gives us Hamiltonian evolution.
 
@@
 \noindent If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.
+This assumption means that momentum can be expressed as a function of position and velocity, which mathematically corresponds to the case in which the Hamiltonian is hyper-regular.
+In this regime we find Lagrangian mechanics.
+Moreover, to be able to express the symplectic form, and therefore the density, over position and velocity, we find that momentum must be a linear function in velocity, which constrains the Hamiltonian to be the one of a massive particle under scalar and vector potential forces.
 
@@
 Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere---which we take to be positive.
+This, by definition, means that motion equivalence implies the Hamiltonian is hyper-regular.
 Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables.

### 2023-02-16 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: minor revisions from Andrea

@@
 Instead of focusing on probability, we use information theoretic considerations to show that all pure states of $\psi$-ontic models must be orthogonal to each other, in clear violation of quantum mechanics.
-Given that  (1) Pusey, Barrett and Rudolph (PBR) previously showed that $\psi$-epistemic models, as defined by HS, also contradict quantum mechanics, and (2) the HS categorization is exhausted by these two types of models, we conclude that the HS categorization itself is problematic as it leaves no space for models that can reproduce quantum theory.
+Given that (i) Pusey, Barrett and Rudolph (PBR) previously showed that $\psi$-epistemic models, as defined by HS, also contradict quantum mechanics, and (ii) the HS categorization is exhausted by these two types of models, we conclude that the HS categorization itself is problematic as it leaves no space for models that can reproduce quantum theory.
 \end{abstract}
@@
 classical) statistical mixture.
-But classical mixtures and quantum mixtures have different properties.
-For example, classical mixtures have a single decomposition in terms of pure states while quantum mixtures do not.
-Studying the behavior of states under mixing, then, provides a new way to explore possible violations.
+However, as is well-known, classical and quantum mixtures have different properties.
+For instance, classical mixtures have a single decomposition in terms of pure states while quantum mixtures do not.
+Studying the behavior of states under mixing, then, provides a new way to explore \textcolor{blue}{possible violations}.
+%of what?
+I would also mention that the example will be relevant for our argument.
 
-Using this new approach, in this paper we are going to provide a no-go theorem for $\psi$-ontic models.
-The core insight is that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.
+\textcolor{blue}{Using this new approach}, %which new approach?
+Io utilizzerei la frase che era presente nella versione precedente: ``using information theoretic considerations we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM since the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all states must be orthogonal, which openly contradicts quantum theory.''
+in this paper we are going to provide a no-go theorem for $\psi$-ontic models.
+\textcolor{red}{The core insight is that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.} %contrary to some previous version of the paper, this sentence is not put in context and some readers not familiar with HS will not understand.
 Consequently, we find that the lack of overlap required by $\psi$-ontic models forces all pure states to be orthogonal.
@@
 
-The exposition of the HS classification makes it seem one has complete freedom in choosing the overlap between epistemic states.
-In this section we will show that this is not the case: this is constrained by entropy of mixtures.
-These constraints cannot be replicated by a $\psi$-ontic model with standard information theoretic techniques.
+%The exposition of the: ci possono obiettare che  la nostra esposizione a non essere fedele, quando invece  proprio il framework di HS d avere il problema di cui parliamo.
+Looking at the HS classification it seems that one has complete freedom in choosing the overlap between epistemic states. %for argumentative strategies I would explain why it seems so.
+In this section we will show that this is not the case: this choice is in fact constrained by entropy of mixtures.
+Moreover, these constraints cannot be replicated by a $\psi$-ontic model with standard (i.e.\ classical) information theoretic techniques.
 Since quantum statistical mechanics and quantum information theory depend on the correct value of entropy on mixtures, $\psi$-ontic models will violate quantum theory.
+%to make our argument stronger we should show that psi-ontic models rely on classical information theoretic structures.
+This fact should clearly stated now not in the next paragraph.
 
@@
 The second will be used to calculate the entropy on the quantum case.
+% always for the sake of clarity, here I would anticipate what we are going to say: given the formal structure on which psi-ontic models are built they cannot reproduce QM.
+In this way the argument will be already "in the head" of our readers.
 
@@
 That is, they are two non-overlapping probability density functions defined to be the Radon-Nikodym derivatives between the respective probability measures over $\Lambda$ and the measure $\lambda$.
-Equation \ref{epistemic_mixing} tells us that the probability measures combine in the usual way, and therefore an equal mixture of the two states must obey equation \ref{entropy_nonoverlap}.
+Equation \ref{epistemic_mixing} tells us that the probability measures combine in a classical way, and therefore an equal mixture of the two states must obey equation \ref{entropy_nonoverlap}.
 Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states in quantum mechanics is zero, we must have
@@
 In other words: \emph{entropy fixes equality}.
-%spiega meglio
 The issue here is that the mapping between epistemic states and density matrices is not injective.
@@
 
-If we combine our result with PBR, the only models that are not ruled out are $\psi$-epistemic models where the objects are not independently prepared.
+If we combine our result with the PBR theorem, the only models that are not ruled out by the HS framework are $\psi$-epistemic models in which quantum objects are not independently prepared.
 That is, where the joint distribution over ontic states is not the product of the marginals.
 However, this loophole can also be closed with information theoretic considerations.
-It is a well-known result in both classical and quantum information theory that the entropy of a joint distribution is the sum of the marginal if and only if the subsystems are independent (\cite{Ash:2010}, \cite{Nielsen:2010}).
+It is a well-known result in both classical and quantum information theory that the entropy of a joint distribution is the sum of the marginal if and only if the subsystems under consideration are independent (\cite{Ash:2010}, \cite{Nielsen:2010}).
 Therefore an epistemic state that represents independent quantum states must also be the joint distribution of statistically independent epistemic states of the individual systems.
@@
 \end{quote}
-Now, the result may seem, at first, preposterous. But once we analyze the mathematics in detail, we see that both results, ours and PBR, are fundamentally finding something, in retrospect, obvious.
+Although this result may at first sight seem paradoxical, once we analyze in detail the mathematics behind HS categorization we see that both results---ours and PBR---are finding something, in retrospect, self-evident.
 
@@
 By using standard information theoretic techniques, we have shown that $\psi$-ontic models are not compatible with quantum information theory, and therefore with quantum mechanics itself.
-Given that the PBR theorem already ruled out $\psi$-epistemic models, the HS categorization of ontological models is fundamentally empty.
+Given that the PBR theorem already ruled out $\psi$-epistemic models, the HS categorization of ontological models is found to be fundamentally empty.
 Thus, such a framework should be employed neither to classify quantum interpretations, nor to draw conclusions on the nature of the quantum state.
@@
 The key problem of the HS categorization is that it is ``not quantum enough.'' The framework tries to hide the non-classicality at the level of each ontic state, which, by construction, cannot be decomposed further using a standard probability measure.
-However, it posits that statistical mixtures are simply classical mixtures in the sense that they are captured by a standard probability measure.
+However, HS definitions entail that statistical mixtures are simply classical mixtures in the sense that they are captured by a standard probability measure.
 But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.
-Therefore the HS categorization prevents one from describing quantum statistical mixing.
+Therefore, the HS categorization prevents one from describing quantum statistical mixing.
 
-Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are not sustainable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions.
-Specifically, the use of standard probability theory in composite systems.
+Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are not sustainable because of an underlying issue of the HS definitions, not because epistemic interpretations contradict quantum theory.
+Specifically, the problem of HS approach is the use of standard probability theory in composite systems.
 If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories, an argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.

### 2023-02-16 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: final pass with everybody

@@
 classical) statistical mixture.
-However, as is well-known, classical and quantum mixtures have different properties.
+However, as it is well-known, classical and quantum mixtures have different properties.
 For instance, classical mixtures have a single decomposition in terms of pure states while quantum mixtures do not.
-Studying the behavior of states under mixing, then, provides a new way to explore \textcolor{blue}{possible violations}.
-%of what?
-I would also mention that the example will be relevant for our argument.
+Using information theoretic considerations on mixtures, then, provides a new approach to explore possible violations of quantum mechanics.
 
-\textcolor{blue}{Using this new approach}, %which new approach?
-Io utilizzerei la frase che era presente nella versione precedente: ``using information theoretic considerations we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM since the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all states must be orthogonal, which openly contradicts quantum theory.''
-in this paper we are going to provide a no-go theorem for $\psi$-ontic models.
-\textcolor{red}{The core insight is that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.} %contrary to some previous version of the paper, this sentence is not put in context and some readers not familiar with HS will not understand.
-Consequently, we find that the lack of overlap required by $\psi$-ontic models forces all pure states to be orthogonal.
-Since this directly contradicts quantum theory, we conclude that $\psi$-ontic models are not viable.
+Using this new approach, in this paper we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM.
+The core insight is that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.
+We find that the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all pure states must be orthogonal, which directly contradicts quantum theory.
 
@@
 
-%The exposition of the: ci possono obiettare che  la nostra esposizione a non essere fedele, quando invece  proprio il framework di HS d avere il problema di cui parliamo.
-Looking at the HS classification it seems that one has complete freedom in choosing the overlap between epistemic states. %for argumentative strategies I would explain why it seems so.
+Looking at the HS classification it may seem that one has complete freedom in choosing the overlap between epistemic states.
 In this section we will show that this is not the case: this choice is in fact constrained by entropy of mixtures.
@@
 Since quantum statistical mechanics and quantum information theory depend on the correct value of entropy on mixtures, $\psi$-ontic models will violate quantum theory.
-%to make our argument stronger we should show that psi-ontic models rely on classical information theoretic structures.
-This fact should clearly stated now not in the next paragraph.
 
-To show the violation, we will first study how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy and the von Neumann entropy.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
+Before proving the violation, we first introduce two results that show how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy and the von Neumann entropy.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
 The second will be used to calculate the entropy on the quantum case.
-% always for the sake of clarity, here I would anticipate what we are going to say: given the formal structure on which psi-ontic models are built they cannot reproduce QM.
-In this way the argument will be already "in the head" of our readers.
 
@@
 That is, they are two non-overlapping probability density functions defined to be the Radon-Nikodym derivatives between the respective probability measures over $\Lambda$ and the measure $\lambda$.
-Equation \ref{epistemic_mixing} tells us that the probability measures combine in a classical way, and therefore an equal mixture of the two states must obey equation \ref{entropy_nonoverlap}.
+Equation \ref{epistemic_mixing} tells us that the probability measures combine in the usual way, and therefore an equal mixture of the two states must obey equation \ref{entropy_nonoverlap}.
 Given that $\psi$ and $\phi$ are pure states, and the entropy for pure states in quantum mechanics is zero, we must have
@@
 
-If we combine our result with the PBR theorem, the only models that are not ruled out by the HS framework are $\psi$-epistemic models in which quantum objects are not independently prepared.
+If we combine our result with the PBR theorem, the only HS ontological models that are not ruled out are $\psi$-epistemic models in which quantum objects are not independently prepared.
 That is, where the joint distribution over ontic states is not the product of the marginals.
@@
 So, how does the HS model try to accommodate this?
-Given that $p(\lambda|P)$ combines using classical probability theory, all the non-classicality must be in $p(k|\lambda, M)$.
+Given that $p(\lambda|P)$ combines using classical probability theory, all the ``quantumness'' must be in $p(k|\lambda, M)$.
 In fact, note that this second expression does not allow us to write the joint probability distribution for multiple observers; therefore, each ontic state is not partitioned further.
-Therefore if $\lambda$ is the state of a single quantum system, we can hide the inner product inside $p(k|\lambda, M)$.
-This is enough to hide our non-classical probability for pure states of single quantum systems, but not much more.
+Therefore, if $\lambda$ is the state of a single quantum system, we can ``hide'' the inner product inside $p(k|\lambda, M)$; this is enough to reproduce quantum probability for pure states of single quantum systems, but not much more.
 
-In our result, we saw that the use of standard measure theory when creating mixtures creates a problem because the convex space of classical distributions is significantly different from the convex space of quantum mixed states.
+In our result, we saw that the use of standard measure theory when creating mixtures as in equation \ref{epistemic_mixing} generates a problem because the convex space of classical distributions is significantly different from the convex space of quantum mixed states.
 For this case, ontological models are ``not quantum enough.''
@@
 \end{equation}
-there is a problem. The issue here is not the factorization. The issue here is breaking the composite system into parts, each with its own ontic state as in 
+there is a problem. However, the issue here is not the factorization, but rather breaking the composite system into parts, each with its own ontic state as in 
 \begin{equation}
@@
 Therefore there are two different ways ontological models use classical measure theory: one in mixtures, highlighted by us; one in composite systems, highlighted by PBR.
-The use of classical measure theory is what drives the failure of the model.
+The use of measure theory is what drives the failure of the model.
 
-The moral of the story is
+The take-home message is
 \begin{quote}
@@
 
-Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are not sustainable because of an underlying issue of the HS definitions, not because epistemic interpretations contradict quantum theory.
-Specifically, the problem of HS approach is the use of standard probability theory in composite systems.
+Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are untenable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions.
+Specifically, the problem of the HS categorization is the use of standard probability theory in composite systems.
 If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories, an argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.

### 2023-02-17 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: minor change for consistency

@@
 
-Therefore there are two different ways ontological models use classical measure theory: one in mixtures, highlighted by us; one in composite systems, highlighted by PBR.
+Therefore there are two different ways ontological models use standard measure theory: one in mixtures, highlighted by us; one in composite systems, highlighted by PBR.
 The use of measure theory is what drives the failure of the model.

### 2023-02-24 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
Update classical-phil-article-Erkenntnis.tex
Edits from GC and JH in response to referee #2, comments 1 thru 4. Major changes in green font for ease of referee and editors. (Josh also changed a few words here and there for style).

--- 2023-02-09 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
@@
 \usepackage{mathptmx}
+
+\usepackage[dvipsnames]{xcolor}%lets us define a new color 
+\definecolor{greeo}{RGB}{91,173,69}
+\colorlet{green}{greeo} %redefines color `green' to color `greeo' (or any other color)
 
@@
 %\author{\Large \textsuperscript{*\textdagger}}
-\date{\vspace{-7.9ex} April 3rd, 2022}
+\date{\vspace{-7.9ex} \today}
 \maketitle
@@
 However, he also shows that if we consider a more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are \textit{not} categorically equivalent \parencites*[1182-83]{Barrett2}.
-Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation.} Whereas North's argument relies on a problematic criterion for comparing structure based on the size of symmetry groups \parencites[]{Swanson}, our argument avoids this pitfall.
+Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation{\color{green} and illuminates why it is more general.}} Whereas North's argument relies on a problematic criterion for comparing structure based on the size of symmetry groups \parencites[]{Swanson}, our argument avoids this pitfall.
 We show that a classical system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian.
 To be Lagrangian, systems must satisfy an additional physical condition, making Lagrangian mechanics a proper subset of Hamiltonian mechanics.
-Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics, and this difference has interpretive consequences even within the hyper-regular domain.
-The construction at the heart of our argument provides an interpretation of classical mechanical systems in terms of their invariant phase space density.
-%We show that the invariant density fits more naturally within the Hamiltonian framework, providing another reason to privilege it over Lagrangian mechanics.
-% Our formalism also provides a novel interpretation of the principle of stationary action within Hamiltonian mechanics.
+Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics.
+{\color{green} Unlike Lagrangian mechanics, Hamiltonian mechanics provides a unique solution to the equations of motion even outside the hyper-regular domain.
+Hence, Lagrangian mechanics is restricted to the hyper-regular regime.} The construction at the heart of our argument provides a novel interpretation of classical mechanical systems in terms of their invariant phase space density.
 
@@
 Section~\ref{derivation} describes in detail the derivation of Hamiltonian mechanics for systems that satisfy two physical conditions: infinitesimal reducibility and deterministic/reversible evolution.
-We then show that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
+{\color{green} Our derivation shows that even non-hyper-regular Hamiltonian systems are physically interpretable.} We then show that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
 This completes our primary argument for privileging Hamiltonian mechanics.
@@
 
-In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must provide a criterion for a framework being more fundamental than another.
-Of various notions of relative fundamentality, we focus on an entirely non-mysterious aspect.
+In claiming that Hamiltonian mechanics is more fundamental than Lagrangian mechanics, we must provide a criterion for assessing whether one framework is more fundamental than another.
+Of various notions of relative fundamentality, we focus on an entirely non-mysterious one.
 Given two formalisms that capture a class of systems, if one is logically more general, then it is more fundamental.
@@
 A more fundamental framework encompasses a less fundamental one as a special case.
-It is in this sense that the generalized Stokes' theorem is more fundamental than well-known special cases, such as the fundamental theorem of calculus and Green's theorem.
-Similarly, an axiomatization which includes additional axioms is less fundamental than one which drops these axioms while maintaining the same expressive power.
-This aspect of fundamentality is distinct from aspects associated with differences in energy scale, such as how the Standard Model of particle physics is more fundamental than classical mechanics.
+It is in this sense that the generalized Stokes' theorem is more fundamental than special cases such as the fundamental theorem of calculus and Green's theorem.
+Similarly, an axiomatization that includes more axioms is less fundamental than one which drops these axioms while maintaining the same expressive power.\footnote{One might worry that our criterion is subject to elementary counterexamples.
+For instance, can't we make any theory more general by adjoining arbitrary disjunctions?
+Importantly, we require that the theory remains wholly \textit{about} classical mechanics.
+Hamiltonian mechanics is more general than Lagrangian mechanics while still being about classical systems, where ``aboutness'' can be understood using Yablo's framework \parencites*[]{Yablo}.
+Having an account of \textit{aboutness} or \textit{subject matter} is a general philosophical problem that any account of fundamentality must address.
+It is not a special problem facing our criterion.}  This aspect of fundamentality is distinct from aspects associated with differences in energy scale, such as how the Standard Model of particle physics is more fundamental than classical mechanics.
 We do not mean that Hamiltonian mechanics is more fundamental than Lagrangian mechanics in this sense.
@@
 We do this in Section~\ref{derivation} by recovering the mathematical formalisms of both Hamiltonian and Lagrangian mechanics from three physical conditions.
-In this section, we describe how our argument works in principle.
-Section~\ref{fundamentality} demonstrates some philosophical advantages of our criterion for relative fundamentality.\footnote{One might worry that our criterion is subject to elementary counterexamples.
-For instance, can't we make any theory more general by adjoining arbitrary disjunctions?
-Importantly, we require that the theory remains wholly \textit{about} classical mechanics.
-Hamiltonian mechanics is more general than Lagrangian mechanics while still being about classical systems, where ``aboutness'' can be understood using Yablo's framework \parencites*[]{Yablo}.
-Having an account of \textit{aboutness} or \textit{subject matter} is a general philosophical problem that any account of fundamentality must address.
-It is not a special problem facing our criterion.}
-%We will show that the Lagrangian framework is less general because it requires an additional condition to hold, compared to the Hamiltonian framework.
-Hence, the Lagrangian framework is less fundamental because it is a proper subset of the Hamiltonian framework.
-
+{\color{green}Our derivation illuminates the relationship between Hamiltonian and Lagrangian mechanics.
+As North \parencites*[]{North} argues, there is a sense in which Hamiltonian mechanics is more general than Lagrangian mechanics.
+Barrett \parencites*[]{Barrett2} makes this sense more precise by showing that the symplectic manifold formulation of Hamiltonian mechanics is categorically inequivalent to Lagrangian mechanics formulated on a tangent bundle.
+Still, one might wonder why it is that Hamiltonian mechanics can be formulated in a more general framework than Lagrangian mechanics.
+Additionally, one might wonder whether Lagrangian mechanics could be meaningfully formulated in an analogously general manner.
+Our derivation helps answer these questions: the physical principles that undergird the symplectic manifold formulation of Hamiltonian mechanics are simply more general than the principles required for Lagrangian mechanics.}\footnote{{\color{green}We thank a referee for this journal for this helpful way of framing the significance of our derivation.}} In this section, we describe how our argument works in principle.
+Section~\ref{fundamentality} demonstrates some philosophical advantages of our criterion for relative fundamentality.
 %We will, however in Section 5 give a criterion for an entity to be physically fundamental
@@
 we cannot specify the whole state by specifying the states of its subsystems).\footnote{The conditions of infinitesimal (ir)reducibility are similar to various notions of (non-)separability in the literature on physical holism \parencites[46, 124]{Healey}.
-We adopt our preferred terminology to avoid controversies surrounding how best to characterize separability.} We show in Section~\ref{infinitesimal} that the phase space of infinitesimally reducible systems must have at least the structure of a symplectic manifold.
-To recover Hamiltonian mechanics, it then suffices to identify when trajectories on phase space obey Hamiltonian evolution (i.e.
+We adopt our preferred terminology to avoid controversies surrounding how best to characterize separability.} We show in Section~\ref{infinitesimal} that {\color{green} if one chooses to model continuous quantities using a manifold, then this manifold must be symplectic.} To recover Hamiltonian mechanics, it then suffices to identify when trajectories on phase space obey Hamiltonian evolution (i.e.
 evolve according to a symplectomorphism).
@@
 We thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution.
-%Since we won't have to settle controversies surrounding these terms, we adopt our preferred terminology.
+{\color{green} Importantly, our derivation encompasses Hamiltonian mechanics in the non-hyper-regular regime as well.}
 
 Next, we note that some classical systems satisfy a third, further condition, which we call \textit{motion equivalence}.
-Motion equivalence holds when the system's trajectory in physical space determines---and is determined by---its evolution in phase space.
+Motion equivalence holds when a system's trajectory in physical space determines---and is determined by---its evolution in phase space.
 In other words, knowledge of the trajectory in physical space (e.g.
@@
 Hence, motion equivalence entails a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics (Section~\ref{motion}).
-As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
+{\color{green} Since a system satisfies motion equivalence if and only if it is hyper-regular, this condition illuminates the physical significance of hyper-regularity.
+Moreover, since Lagrangians lack unique solutions to the equations of motion outside the hyper-regular regime, we take ourselves to have recovered the physically significant Lagrangian systems.} As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
 
@@
 For at least this reason, Hamiltonian mechanics is more fundamental than Lagrangian mechanics.
-
-% % optional: address refer 1 concern about the clarity of this argument/example: ``what if momentum is defined in terms of velocity? Or is the idea to know that we also need to know the mass?"
 
@@
 This supports our claim that the property of infinitesimal reducibility fruitfully characterizes classical systems.
-Before presenting the details of our argument in Section~\ref{derivation}, we describe how we avoid central problems facing North's privileging of Hamiltonian mechanics.
+Before presenting the details of our argument in Section~\ref{derivation}, we describe how we avoid central problems that arise in North's privileging of Hamiltonian mechanics.
 %, although Curiel seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
@@
 Our derivation has three main steps.
-Section~\ref{infinitesimal} shows that the phase space of infinitesimally reducible systems must have at least the structure of a symplectic manifold.
-For systems with deterministic and reversible evolution, Section~\ref{deterministic} shows they must evolve according to a symplectomorphism.
+Section~\ref{infinitesimal} shows that {\color{green} if one represents continuous quantities using a manifold, then this manifold must be symplectic.} For systems with deterministic and reversible evolution, Section~\ref{deterministic} shows they must evolve according to a symplectomorphism.
 Hence, we derive the formalism of Hamiltonian mechanics from the conditions of infinitesimal reducibility and deterministic/reversible evolution.
 Section~\ref{motion} shows how systems that satisfy the further condition of motion equivalence can be treated within the formalism of Lagrangian mechanics.
+{\color{green}In fact, when motion equivalence fails, Lagrangian mechanics does not yield unique solutions to the equations of motion.}
 
@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.}
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.} {\color{green} Ultimately, this condition entails that we can describe the state of the whole system as a statistical distribution over the state space of its parts.
+If we model the state space of the parts using a manifold, this manifold must be differentiable so that it can support a density at each point.
+Moreover, given that this density must be a function of the state, and not of the coordinates, we show that the manifold must be symplectic.}
+% % Could add: readers who are satisfied with this overview can proceed to the next subsection
 % % % For further discussion, see \textcites{shpmp}. [omit this for review].
@@
 
-When the system is characterized by continuous quantities, $\mathcal{S}$ is a manifold.
-For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
+When the system is characterized by continuous quantities, {\color{green}$\mathcal{S}$ can be taken to be a manifold}.\footnote{{\color{green}The use of manifolds is, technically, a choice of how the physics is represented.
+One could make another choice, for example, by representing the state variables not as local coordinates but as an algebra of operators that act on the state.
+In this case, we conjecture that a similar argument would demonstrate an analogue of a symplectic structure for the different mathematical representation.}} For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
 \textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called ``coordinates'' and $\xi$ is a coordinate map.
@@
 \noindent In our case, this applies to both the state of the whole system and the state of its parts.
+{\color{green} Under this assumption, we show that the symplectic structure must be preserved over time, which gives us Hamiltonian evolution.}
 
@@
 In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.
+{\color{green} Note in particular that our derivation is not restricted to only hyper-regular Hamiltonians.
+As we will see, this is a key difference between Hamiltonian and Lagrangian mechanics.}
+
+%Since, the only requirement that we have imposed on the Hamiltonian is that the equations yield a unique solution, our derivation applies to Hamiltonians that are not hyper-regular. }
+
+% since we have not imposed hyper-regularity,   }
 % [(ref 1): expand on how our argument relates to standard proofs of Liouville's theorem.
@@
 \noindent If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.
+{\color{green}This assumption means that momentum can be expressed as a function of position and velocity, which mathematically corresponds to the case in which the Hamiltonian is hyper-regular.
+In this regime we find Lagrangian mechanics.
+Moreover, to be able to express the symplectic form, and therefore the density, over position and velocity, we find that momentum must be a linear function in velocity, which constrains the Hamiltonian to be the one of a massive particle under scalar and vector potential forces.}
 
@@
 Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere---which we take to be positive.
-Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables.
+{\color{green} This, by definition, means that the Hamiltonian is hyper-regular if and only if weak equivalence is satisfied.} Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables.
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).

### 2023-03-03 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
Update classical-phil-article-Erkenntnis.tex
Some more minor changes, including citation to Gelfand and Fomin

@@
   Pagetotal     = {xiii, 130 P.},
+}
+
+@Book{Gelfand,
+  Title     = {Calculus of Variations},
+  Publisher = {Prentice--Hall},
+  Year      = {1963},
+  Author    = {Gelfand, I. M. and Fomin, S. V.},
+  Address   = {Englewood Cliffs, N.J.},
+  translator = {Silverman, Richard A.},
+  Note      = {revised english edition; Silverman edited it as well and added problems},
 }
@@
 However, he also shows that if we consider a more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are \textit{not} categorically equivalent \parencites*[1182-83]{Barrett2}.
-Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation{\color{green} and illuminates why it is more general.}} Whereas North's argument relies on a problematic criterion for comparing structure based on the size of symmetry groups \parencites[]{Swanson}, our argument avoids this pitfall.
+Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation {\color{green} and illuminates why it is more general.}} Whereas North's argument relies on a problematic criterion for comparing structure based on the size of symmetry groups \parencites[]{Swanson}, our argument avoids this pitfall.
 We show that a classical system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian.
@@
 Section~\ref{Lagrangian} considers the status of Lagrangian mechanics within our framework.
-We provide a geometric interpretation of the action principle, which applies even outside the hyper-regular domain.
+We provide a geometric interpretation of the action principle that applies even outside the hyper-regular domain.
 We end by rebutting objections that arise from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
@@
 A more fundamental framework encompasses a less fundamental one as a special case.
-It is in this sense that the generalized Stokes' theorem is more fundamental than special cases such as the fundamental theorem of calculus and Green's theorem.
-Similarly, an axiomatization that includes more axioms is less fundamental than one which drops these axioms while maintaining the same expressive power.\footnote{One might worry that our criterion is subject to elementary counterexamples.
+It is in this sense that the generalized Stokes' theorem is more fundamental than special cases such as the fundamental theorem of calculus and Green's theorem.\footnote{One might worry that our criterion is subject to elementary counterexamples.
 For instance, can't we make any theory more general by adjoining arbitrary disjunctions?
@@
 As Gilton notes, ``many philosophers of science have moved away from thinking of fundamentality in science in terms of the very small toward thinking of it in terms of comprehensiveness, or breadth of applicability'' \parencites*[637]{Gilton2021}.
+% % Removed: Similarly, an axiomatization that includes more axioms is less fundamental than one which drops these axioms while maintaining the same expressive power.
+%Worry: what if  both axiom systems cover the same class of systems? then seemingly no difference in generality? 
 %we must provide a criterion for one framework being more or less fundamental than another
@@
 Additionally, one might wonder whether Lagrangian mechanics could be meaningfully formulated in an analogously general manner.
-Our derivation helps answer these questions: the physical principles that undergird the symplectic manifold formulation of Hamiltonian mechanics are simply more general than the principles required for Lagrangian mechanics.}\footnote{{\color{green}We thank a referee for this journal for this helpful way of framing the significance of our derivation.}} In this section, we describe how our argument works in principle.
+Our derivation helps answer these questions: the physical principles that undergird the symplectic manifold formulation of Hamiltonian mechanics are simply more general than the principles required for Lagrangian mechanics.}\footnote{{\color{green}We thank a referee for suggesting this helpful way of framing the significance of our derivation.}} In this section, we describe how our argument works in principle.
 Section~\ref{fundamentality} demonstrates some philosophical advantages of our criterion for relative fundamentality.
@@
 {\color{green} Since a system satisfies motion equivalence if and only if it is hyper-regular, this condition illuminates the physical significance of hyper-regularity.
-Moreover, since Lagrangians lack unique solutions to the equations of motion outside the hyper-regular regime, we take ourselves to have recovered the physically significant Lagrangian systems.} As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
+Moreover, since Lagrangians lack unique solutions to the equations of motion outside the hyper-regular regime, we take ourselves to have recovered the physically significant Lagrangian systems.}\footnote{{\color{green}See \textcites[119]{Gelfand} for discussion of the fact that Lagrangians yield unique solutions only in the hyper-regular domain.}} As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
 
@@
 Of course, different definitions of ``Hamiltonian'' and ``Lagrangian'' systems can result in different interpretive relationships.
-On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems also satisfy the motion equivalence assumption.
+On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems also satisfy the motion equivalence condition.
 Since Lagrangian systems satisfy all three conditions (infinitesimal reducibility, deterministic/reversible evolution, and motion equivalence), they lie in the intersection of Newtonian and Hamiltonian systems.} We thereby set aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \textcites*[3-4]{Baez}{Earman}{Norton}.} Additionally, we set aside dissipative systems, i.e.
@@
 Curiel \parencites*[311]{Curiel} acknowledges that the Hamiltonian framework can be extended to treat dissipative systems.
-For details of a Lagrangian approach to dissipative systems, see \textcites[]{Smith}.} This restricts our claim about the fundamentality of Hamiltonian mechanics to the non-dissipative regime.
+For a Lagrangian approach to dissipative systems, see \textcites[]{Smith}.} This restricts our claim about the fundamentality of Hamiltonian mechanics to the non-dissipative regime.
 Finally, our treatment is not restricted to $n$-particle mechanics but applies as well to rigid-body and continuum mechanics.
@@
 Compared with North's \parencites*[]{North} account of comparative structure, our criterion has two advantages.
-First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which scientific anti-realists and pragmatists view with suspicion.
+First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which scientific anti-realists, pragmatists, and even many naturalists view with suspicion.
 Secondly, it avoids a technical objection that Swanson and Halvorson \parencites*[]{Swanson} levy against North's more precise mathematical criterion for comparing structure.
+% Could cite Woodward, or Cohen/Callender for skepticism about perfectly natural properties. Includes naturalistically-minded philosophers
 %provided a theoretical framework captures the relevant behavior, it is more fundamental if it requires fewer conditions to hold.
@@
 Owing to the epistemic difficulty of accessing these putative joints, we do not endorse these metaphysical dimensions of structure or fundamentality.\footnote{For a detailed discussion of difficulties facing the notion of metaphysical naturalness, see \textcites[]{Dorr_Hawthorne}.} While Hamiltonian mechanics is significant partly because it elucidates invariant densities, this does not entail that additional mathematical structure is physically insignificant or surplus.
-Additional structure may correspond to an additional physically-significant assumption that applies to the system under consideration.
+Additional structure may correspond to an additional physically-significant condition that applies to the system under consideration.
 Precisely this scenario occurs in the context of Lagrangian mechanics, which requires the additional condition of motion equivalence to hold.
-Whereas North views Lagrangian mechanics as containing ``excess, superfluous structure'' \parencites*[75]{North}, we take this additional condition as optional but still physically significant.
+Whereas North views Lagrangian mechanics as containing ``excess, superfluous structure'' \parencites*[75]{North}, we take this additional condition to still be physically significant.
 Thus, while it is illuminating to identify the minimum mathematical structures necessary for a formulation, additional structure is not automatically rendered physically insignificant.
 
-Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real'' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret statespace structure ontologically, but as a formal device for encoding physical information or a calculational tool.
+Additionally, whereas North \parencites*[76]{North} seeks to determine the ``real'' fundamental state space structure of the theory---and correspondingly, of a classical world---we remain neutral on this underlying metaphysical question.\footnote{Indeed, Swanson and Halvorson \parencites*[]{Swanson} note that many physicists and philosophers would not interpret state-space structure ontologically, but rather as a formal device for encoding physical information or a calculational tool.
 As Wilson illustrates, we can interpret classical mechanics while avoiding---and even rejecting---the search for what is ``fundamental in the bottom-layer sense'' \parencites*[53]{Wilson}.} Whether Hamiltonian phase space structure is metaphysically fundamental in a classical world does not matter for the questions we seek to answer.
@@
 
-We use the three physical conditions to place constraints on both physical entities and, in turn, the mathematical objects that can be used to represent them.
+These three physical conditions place constraints on both physical systems and, in turn, the mathematical objects that can be used to represent them.
 Each condition leads to additional constraints, introduced gradually within each section.
@@
 
-When the system is characterized by continuous quantities, {\color{green}$\mathcal{S}$ can be taken to be a manifold}.\footnote{{\color{green}The use of manifolds is, technically, a choice of how the physics is represented.
+When the system is characterized by continuous quantities, {\color{green}$\mathcal{S}$ can be taken to be a manifold}.\footnote{{\color{green}Note that the use of manifolds constitutes a choice of how to represent the physics.
 One could make another choice, for example, by representing the state variables not as local coordinates but as an algebra of operators that act on the state.
-In this case, we conjecture that a similar argument would demonstrate an analogue of a symplectic structure for the different mathematical representation.}} For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
+In this case, we conjecture that a similar argument would demonstrate an analogue of a symplectic structure for this alternative mathematical representation.}} For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
 \textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called ``coordinates'' and $\xi$ is a coordinate map.
@@
 It allows us to compare and track the amount of matter in different possible states.
+%The use of manifolds is, technically, a choice of how the physics is represented
 
@@
 
-Moreover, the units we choose for $\omega$ are the units we choose to count states over an independent degree of freedom.
-This agrees with the practice in statistical mechanics.
+Moreover, we choose the same units for $\omega$ as the units we choose to count states over an independent degree of freedom.
+This matches common practice in statistical mechanics.
 When normalized, the density $\rho_{c, \xi}$ is given in terms of the inverse of the units for $\omega$.
-In accordance with the standard conventions in quantum mechanics and statistical mechanics, the $k_i$ are given in terms of inverse units of the respective $q^i$, such that $dq^i \wedge dk_i$ is a pure number.
+Following standard conventions in quantum mechanics and statistical mechanics, the $k_i$ are given in terms of inverse units of the respective $q^i$, such that $dq^i \wedge dk_i$ is a pure number.
 We then use $\hbar$ to define the units for state counting and therefore set $\omega = \hbar dq^i \wedge dk_i = dq^i \wedge dp_i$ where $p_i = \hbar k_i$.
+
+%Moreover, the units we choose for $\omega$ are the units we choose to count states over an independent degree of freedom. This agrees with the practice in statistical mechanics.
 
@@
 trajectories in physical space) suffices to recover its dynamics (i.e.
-evolutions in state space) and vice-versa.\footnote{ \textcites[218]{Abraham}, Theorem 3.6.2 shows that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time; see \textcites[1180-1181]{Barrett2} for discussion.
+evolutions in state space) and vice-versa.\footnote{ \textcites[218]{Abraham} show that in the hyper-regular domain, Lagrangian and Hamiltonian mechanics describe the same evolution of particle positions over time (Theorem 3.6.2).
+See \textcites[1180-1181]{Barrett2} for discussion.
 Motion equivalence can be understood as the converse of this theorem: assuming that Lagrangian and Hamiltonian mechanics agree on particle evolutions, we derive that the Legendre transformation is well-defined in the hyper-regular domain.} \es
@@
 In this regime we find Lagrangian mechanics.
-Moreover, to be able to express the symplectic form, and therefore the density, over position and velocity, we find that momentum must be a linear function in velocity, which constrains the Hamiltonian to be the one of a massive particle under scalar and vector potential forces.}
+Moreover, to be able to express the symplectic form (and therefore the density) over position and velocity, we find that momentum must be a linear function in velocity, which constrains the Hamiltonian to be that of a massive particle under scalar and vector potential forces.}
 

### 2023-03-06 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
ClassMechPhil: minor change

@@
 We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.} {\color{green} Ultimately, this condition entails that we can describe the state of the whole system as a statistical distribution over the state space of its parts.
-If we model the state space of the parts using a manifold, this manifold must be differentiable so that it can support a density at each point.
-Moreover, given that this density must be a function of the state, and not of the coordinates, we show that the manifold must be symplectic.}
+If we model the state space of the parts using a manifold, the distribution must be characterized by a density that depends on the state, and not on the coordinates, which leads us to find that the manifold must be symplectic.}
 % % Could add: readers who are satisfied with this overview can proceed to the next subsection

### 2023-03-06 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
Version resubmitted and Response to Referees
version resubmitted to Erkenntnis!

@@
 
-\title{\vspace{-3em} \LARGE Hamiltonian Privilege}			
+\title{\vspace{-4em} \LARGE Hamiltonian Privilege}			
 %\author{\Large \textsuperscript{*\textdagger}}
@@
 On our account, Hamiltonian systems are infinitesimally reducible and follow deterministic and reversible evolution.\footnote{Although not sufficient for a definition, Newtonian systems also satisfy the motion equivalence condition.
-Since Lagrangian systems satisfy all three conditions (infinitesimal reducibility, deterministic/reversible evolution, and motion equivalence), they lie in the intersection of Newtonian and Hamiltonian systems.} We thereby set aside well-known cases of non-deterministic classical systems.\footnote{For discussion of non-deterministic classical systems, see \textcites*[3-4]{Baez}{Earman}{Norton}.} Additionally, we set aside dissipative systems, i.e.
+Since Lagrangian systems satisfy all three conditions (infinitesimal reducibility, deterministic/reversible evolution, and motion equivalence), they lie in the intersection of Newtonian and Hamiltonian systems.} We thereby set aside well-known cases of non-deterministic classical systems.\footnote{For non-deterministic classical systems, see \textcites*[3-4]{Baez}{Earman}{Norton}.} Additionally, we set aside dissipative systems, i.e.
 those that fail to conserve energy.
@@
 This supports our claim that the property of infinitesimal reducibility fruitfully characterizes classical systems.
-Before presenting the details of our argument in Section~\ref{derivation}, we describe how we avoid central problems that arise in North's privileging of Hamiltonian mechanics.
+Before presenting the details of our argument in Section~\ref{derivation}, we describe how we avoid problems that arise in North's privileging of Hamiltonian mechanics.
 %, although Curiel seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
@@
 
-First, we will show that a classical phase space arises if a system satisfies the following physical condition
+First, we will show that a classical phase space arises if a system satisfies the following physical condition:
+\newpage % % %%remember to take out!
 
@@
 We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.} {\color{green} Ultimately, this condition entails that we can describe the state of the whole system as a statistical distribution over the state space of its parts.
-If we model the state space of the parts using a manifold, the distribution must be characterized by a density that depends on the state, and not on the coordinates, which leads us to find that the manifold must be symplectic.}
+If we model the state space of the parts using a manifold, the distribution must be characterized by a density that depends on the state and not on the coordinates.
+From this condition, we show that the manifold must be symplectic.}
 % % Could add: readers who are satisfied with this overview can proceed to the next subsection
@@
 The derivation uses the extended phase space $(q, p, t)$, i.e.
-the standard phase space extended by the time variable.
+the standard phase space extended by a time variable.
 Knowing that the system is deterministic and reversible, it follows from Section~\ref{deterministic} that the density of states is invariant.
@@
 
-By definition, a variation over the integral $\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ consists in taking the difference between the integrand and a corresponding integral over a nearby trajectory $\gamma'$: $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} - \integral_{\gamma'} \vector{\theta} \cdot d\vector{\gamma'}$ .
+By definition, a variation over the integral $\integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ consists in taking the difference between the integrand and an integral over a nearby trajectory $\gamma'$: $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} - \integral_{\gamma'} \vector{\theta} \cdot d\vector{\gamma'}$ .
 Since the trajectories $\gamma$ and $\gamma'$ have the same endpoints, they define a closed-curve $C$, enclosing a region $\Sigma $.
 Hence, by Stokes' theorem $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma} = \oint_{C} \vector{\theta}  \cdot \vector{dC} = \iint_{\Sigma} curl(\vector{\theta}) \cdot \diff \vector{\Sigma}$ .
-Recalling that i) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and ii)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $curl(\vector{\theta})$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, it follows that $\vector{S}$ is tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
+Recalling that i) the $curl(\vector{\theta})$ equals the negative of the displacement vector $\vector{S}$, and ii)  $\vector{S}$ is tangent to the trajectory $\gamma $, we see that $curl(\vector{\theta})$ is tangent to the surface $\Sigma $.\footnote{Since $\vector{S}$ is tangent to $\gamma $, and $\gamma'$ defines an infinitesimal variation, $\vector{S}$ must be tangent to the entire surface $\Sigma $.} Hence, the surface integral must vanish, showing that $\delta \integral_{\gamma} \vector{\theta} \cdot d\vector{\gamma}$ does indeed equal zero.
 This establishes a more general version of Hamilton's principle of stationary action, which we have derived simply by assuming that the displacement vector is divergenceless.
@@
 Instead, it arises in all contexts where motion equivalence holds, i.e.
-where the trajectory of the system in physical space is necessary and sufficient for its evolution in state space.
+where knowledge of the trajectory of the system in physical space is necessary and sufficient for knowledge of its evolution in state space.
 Thus, we take ourselves to have responded to Curiel's main criticism of Hamiltonian mechanics, namely that it must impose \textit{ad hoc} the condition (\ref{6.2}).
@@
 
-Whereas we can easily interpret non-hyper-regular Hamiltonians, the same cannot be said for non-hyper-regular Lagrangians.
-For example, consider the case where the Lagrangian $L$ equals zero.
+Whereas we can easily interpret non-hyper-regular Hamiltonians, non-hyper-regular Lagrangians pose interpretative challenges.
+For example, consider the case where the Lagrangian equals zero.
 For this ``system,'' all paths have the same action, so all paths are possible.
@@
 Indeed, Curiel acknowledges that only hyper-regular Lagrangians have unique solutions whereas Hamiltonian systems have unique solutions even outside the hyper-regular domain \parencites*[291, 308]{Curiel}.
-Curiel argues this favors Lagrangian mechanics, but it generates challenges for understanding non-hyper-regular Lagrangians.
+Although Curiel argues this favors Lagrangian mechanics, it generates challenges for understanding non-hyper-regular Lagrangians.
 %\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.

### 2023-04-13 Articles/2023-OperationalTime/OperationalTime.tex
OperationalTime: initial draft

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amsfonts}

\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\begin{document}

\title{Modeling time from operational requirements}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
In a previous work, We present a theoretical framework in which the adage ``time is what the clock measures'' can be made mathematically precise. The core idea is not to focus on the processes that one uses to implement a clock, but rather on the requirements that such an object must have. We model a clock as a set of references that can be used to tell whether an event is before or after each reference. We then find the necessary and sufficient conditions for a set of references to identify the points of a real line, thus finding what are the physical assumptions we are implicitly taking when modeling time as a continuous quantity. We will find these conditions to be highly idealized and, that once those idealizations are no longer tenable, the linear structure of time breaks done. Under those circumstances the standard notion of manifold, real number or integer need to be abandoned in favor of a yet to be discovered topological structure.
\end{abstract}

\maketitle

\section{Introduction}

The mathematical structure of time (i.e. real numbers with the standard topology) can be understood as coming from an idealized operational model of clocks

As time resolution increases, this model fails in such a way that standard mathematical tools (e.g. manifolds, real numbers, rational numbers or integers) must be abandoned.

A better understanding of how time and space break down at Plank scale, then, means modeling the conceptual issues stemming from less idealized operational definitions.

\section{Clocks, time and synchronization}

The first thing we should be aware is that there are always multiple ways to define a concept, and whether a particular definition is appropriate depends on context. For example, a table made of gold and diamonds may be considered furniture for the purpose of interior decorating, though it may not for insurance or tax purposes. Moreover, if one defines a table as ``a piece of furniture consisting of a smooth flat slab fixed on legs''\footnote{Merriam-Webster definition - https://www.merriam-webster.com/dictionary/table}, there is an implicit understanding of what is meant by ``legs''. Therefore a definition of a concept cannot be given unless we specify its purpose and what are the primitive elements allowed. The question ``what is time?'' is unlikely to have a single answer, not because of the nature of ``time'', but because of the nature of definitions.

For example, in mathematics the main goal of a definition is formal rigor, and the primitive elements are typically those built on top of propositional calculus and set theory. A definition of time may sound like ``a variable that can be used as a parameter for the evolution''. That is, it is the symbol $t$ we use to indicate the argument of the evolution $x^i(t)$, which is a function that takes a real value and returns a point on a manifold. On the other hand, if one is interested in metaphysics, he will gravitate more towards something like ``a nonspatial continuum in which events occur''. If the interest is epistemology, other features will be preferred.

For experimental physics, the goal is to be able to perform experiments, therefore the most important feature is that it gives us a procedure that anyone can follow and gives consistent results. In that context ``time is what a clock measures''. That is, we must have an instrument that measures time, and we must be able to read its output to record the moment an event occurs. In this context, a formal or an ontological definition does not help, and therefore they are not useful, much in the same way being able to perform open heart surgery would not help in the cockpit of a commercial airplane.

While the above definition gives an operational way to define time (i.e. go read the clock), we are left with the question ``what is a clock?'' The obvious answer, ``a clock is what measures time'', fails because it doesn't tell us how to procure a clock. Historically, there have been many objects that have been used as a clock: the motion of the sun as measured by a sundial, the beating of one's heart felt by the pulse, the level of the water of a bucket with a hole, the swings of a pendulum, and so on. For all these devices we can produce a set of instructions such that anybody can construct their own clock, or buy it from someone else.

What we have done, however, is simply give examples of clocks. We haven't answered the question ``what is a clock?" Suppose we had a new device in front of us, how could we tell whether it can be used as a clock? Operationally, if it produces a signal that can be related with the other clocks that would be enough. For example, if it made a chirping noise every so often, and we noticed that it chirped twenty-three times every four swings of a pendulum, or every time the level of the water in our bucket went down thirteen millimeters, then we could use it as a clock. That is, a clock is any device that can be synchronized with other clocks. By synchronization here we mean the ability to relate the measurements of the two clocks, which is another operational feature.

Now, the mathematician may object that these procedures are not ``rigorous'' enough. The metaphysician may object that we made no progress in saying what time ``is''. The epistemologist may object that we shouldn't extrapolate the future behavior of clocks solely based on their past behavior. All of that may be true, but we are able to create clocks and measure time, which is what we needed to achieve for experimental science. The remaining question is: ``are the above considerations enough for theoretical physics?'' That is, should theoretical physics only concern itself with operational definitions, or is there is something more?

To answer this question, instead of exposing a point of view or providing arguments, we use mathematical modeling. We identify the operational requirements and assumptions needed to model clocks and turn them into precise mathematical definitions. If we are able to rigorously recover how time is represented in physics, that is, as a real valued quantity with the standard topology, together with a physically meaningful characterization of each mathematical concept, then we know that all that theoretical physics is describing are the relationships given by the (perhaps idealized) operational definitions.

If we are able to do this, then we should be extremely cautious in using physics to infer anything beyond the operational settings. On the other hand, it would tell us that to improve our understanding of time we should not study the current theoretical frameworks, but rather the idealizations that we are, maybe unknowingly, assuming when devising and using our measurement devices.

\section{The logic of experimental verifiability}

We have clarified that the purpose of our definitions is to capture operational procedures, so now we need to understand how to capture them in a formal setting. The general idea is that we need to capture the logic of experimentally verifiable statements, which means capturing the rules used to combine not only logical statements, but experimental tests as well.\footnote{TODO: reference book and Kevin Kelly}

Our most basic notion is that of a \textbf{verifiable statement}. That is, a statement that is universally true or false for which we have available an experimental test that always terminates successfully in a finite time if and only if the statement is true. For example, ``water is transparent\footnote{At visible wavelengths}'' is a verifiable statement as we can find some and confirm that we can see through it. By statement we mean the assertion made by the sentence, not the sentence itself. Therefore ``l'acqua \`e trasparente'' is the same statement as before, as it is just expressed in another language.

Not all statements are verifiable. Some, like ``chocolate is good'', are not universal; others, like ``23 is a prime number'', ``it is morally justifiable to kill a person to save ten others'' or ``God is eternal'', deal with concepts that cannot be defined operationally; others, like ``the mass of the electron is exactly 0.511 $MeV/c^2$'', are not experimentally verifiable as they would require infinite precision. However, ``the mass of the electron is not exactly 0.511 $MeV/c^2$'' is verifiable, as finite precision measurements can exclude values.

Note that the specific meaning of each statement and its logical relationship with respect to the others cannot be given in isolation. For example, ``the mass of the electron is within (0.510, 0.511) $MeV/c^2$'' is a testable hypothesis if we are trying to measure the mass. However, if we are performing particle identification in a collider experiment, the statement is assumed to be true. Even a physical system, like an electron, may have different operational definitions (e.g. an element of the cathode ray, a $\beta$ particle from radioactive decay) that in some experimental setups need to be shown to be equivalent, while in others are simply assumed to be equivalent based on the other experiment. Therefore verifiable statements must live in a \textbf{logical context} that fully determines these relationships and that is, in general, disconnected from other contexts.

It should be intuitive to anyone familiar with the practice of experimental science that a complete and formal specification of a verifiable statement, together with its test and experimental setup is neither required nor desirable. In the end, the level of detail required is just so that an expert of the field is able to replicate the setup, which allows one to forego a lot of details that would make the specification unnecessarily cumbersome. Moreover, the flexibility to make modifications is essential to customize the design to different settings (i.e. different parts available, different laboratory conditions, etc...). Therefore trying to specify operational procedures themselves in a formal language is unnecessary and impracticable. We are also skeptical that one would be able to come up with a language that would work for all past and future measurement setups.\footnote{One can also argue that defining a language that defines all and only the operational procedures that are possible is equivalent to providing the laws of physics, which makes the problem circular: to define the laws of physics we need an operational language that needs the laws of physics to be defined.} Therefore we avoid this entirely and consider verifiable statements to be primitives. In the same way that mathematical axioms do not specify what sets are, just their properties in terms of set memberships and allowed operations, our formal system will just assume that these statements exist and that they have some basic properties.

Formally, we have a set, called logical context, which contains elements that are called statements. All logical properties between statements are defined in terms of the possible assignments, which assign truth values to all statements in a way that is consistent with their meaning. For example, in all possible assignments ``animal A is a cat'' and ``animal A is a dog'' are never both true. Logical operations are defined as maps within the context. For example, the conjunction is a map that takes two or more statements and returns another statement that is true only for those possible assignments in which all initial statements are true. One finds, as expected, that the logical context is a complete Boolean algebra. What is important to us is that the algebra itself does not depend on what is factually true or not, just what can be potentially true based on the assumptions and operational requirements defined by the context.

Within each context, there is a subset of statements that are verifiable which obey some additional rules. For example, the finite conjunction and countable disjunction of verifiable statements is also a verifiable statement. These rules are justified by the ability to construct tests from other tests. For example, to test a finite conjunction we can run all the tests and, if the statement is true, all tests will finish successfully in finite time. For a countable conjunction, we only need one test to successfully terminate, which we can find within a countable number of tests. These considerations make the set of verifiable statements a Heyting algebra, and the interplay of the Heyting algebra of verifiable statements with the Boolean algebra of all statements gives us a rich mathematical structure.

Even given an indefinite amount of time, we can test up to countably many verifiable statements. Therefore we define an \textbf{experimental domain} as a set of verifiable statements that can be generated by countably many statements through finite conjuction and countable disjunction. Conceptually, an experimental domain corresponds to all the information that can be acquired experimentally on a subject. We then define the \textbf{theoretical domain} as the closure of an experimental domain by negation and countable disjuction. This corresponds to all statements that can be assigned an experimental test, with no guarantee of termination. One can show that the theoretical domain is an atomic lattice, and we call \textbf{possibilities} the atoms. These are the narrowest, most specific, statements and correspond to the cases that can be experimentally distinguished given the verifiable statements of the experimental domain.

This construction maps perfectly to topologies and $\sigma$-algebras in the following way:
\begin{itemize}
	\item possibilities (experimentally distinguishable cases) $\to$ set of points
	\item experimental domain (verifiable statements) $\to$ topology (open sets)
	\item theoretical domain (statements with a test) $\to$ $\sigma$-algebra (Borel sets).
\end{itemize}

Over a real line, then, the possibilities correspond to the points, to all possible values. The standard topology corresponds to assuming that all open intervals correspond to physically verifiable statements. That is, every measurement has, at best, finite precision. Singletons, sets with only one element, are not open sets but are closed set. Physically, we cannot verify a value with infinite precision, but we can falsify it. This establishes a direct link between topological structures and experimental verification.

We are now in a position to formulate our original question precisely both conceptually and mathematically. Each clock will, ultimately, provide a set of verifiable statements (e.g. ``the event happened before clock A ticked 35 times''). All these statements, taken together, will form an experimental domain. What are the necessary and sufficient conditions on all the verifiable statements constructed from all the clocks such that time can be modeled by a real line with the standard topology?

\section{References and ordering}

We model clocks as a set of references, and each reference allows us to verify whether an event happened before or after the reference. A reference, is formalized as a triplet of statements, before/on/after, with before and after guaranteed to be verifiable. In general, both a reference and an event will have a finite duration in time and they may overlap in different ways. [TODO add pictures]. This means that before/on/after are not, in general mutually exclusive. However, we do have some constraints: an event must happen, therefore at least one of the three statements must be true; if the event happens before and after, then it must also happen on the reference.

The main technical result in [reference book] shows the necessary and sufficient conditions for which the possibilities of the domain are ordered, and the topology corresponds to the order topology. The difficult part is not finding the conditions for continuous time but for ordered time. There are three conditions.

\textbf{Strict references.} This tells us that, for each reference, the before/on/after statements must be mutually incompatible: exactly one of then must be true. Conceptually this would mean, for example, that the event we are identifying has an extent in time that is much smaller than the extent of the reference, and therefore the overlaps between event and reference can be neglected.

\textbf{Aligned references.} Formally, this imposes that the before and after statements of a set of references are aligned by implication (e.g. if the event is before reference A then it is not after reference B; if it is before reference C then it is before reference D). Physically, to be able to guarantee precise before/after relationships between references we must have control of the jitter. If there is jitter, then the ordering between references across clocks may not be reliable enough to assume aligned references. That is, the references of two different clocks may shift in opposite directions which may affect their ordering.

\textbf{Refinable references.} This means that we can refine the range by finding references that resolve overlaps or intervals between references. For example, if an event can be after reference A and before reference B, then we must be able to put a reference in between. Conversely, if different references overlap, we must be able to find a reference within that overlap.

When and only when all these conditions are met, then we can assume that the property identified by the references is an ordered property, a quantity. The values, then, correspond to references that are precisely only on the corresponding value. The difference between a continuous quantity, modeled with a real value, and a discrete quantity, modeled with an integer, is whether between two references there is always a third. Therefore the references become infinitely small, with zero extent.

Again we stress that the conditions are simply in terms of logical relationships between verifiable statements, between operational procedures: if this clock tells me that the event happened before the 30 minutes mark, then this other clock must tell me that the event happened before the 35 minutes mark. The open sets, the verifiable statements, correspond to finite precision intervals, and the mathematical structure maps fully to formalized operational requirements. There is nothing else.

The same derivation works for all other quantities, including spatial quantities. Therefore using real valued quantities and manifold in physics is implicitly assuming that, operationally, the properties we are measuring can be distinguished with strict, aligned and refinable references. The theoretical underpinning of this particular choice of mathematical representation are fully captured by operational requirements. In light of this, we should be extremely skeptical of deriving any other consequence from this choice.

\section{Ordering failure}

Assumptions of the order theorem are untenable. Inevitable failure of ordering. Time is cannot be modeled by a continuous or a discrete quantity.

The need for a new mathematical structure. Ordered at large scale but not at short scale. Possible failure of the division of topology/geometry.

The possibility of space-time contextuality.

\section{Conclusion}

\bibliography{bibliography}

\end{document}

### 2023-05-16 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: Adding last paragraph

@@
 
+As a final note, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models.
+To our knowledge, however, none of them is able to reproduce all results from quantum statistical mechanics (e.g.
+derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g.
+derive the Holevo bound).
+Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.
+
 \section{Acknowledgements}

### 2023-05-19 Articles/2021-WhyClassicalMechanicsFails/WhyClassicalMechanicsFails.tex
Fixed typos

@@
 
-Before we explore more what happens across the system-environment boundary, we note that quantum mechanics does suffer from the same problem.
+Before we explore more what happens across the system-environment boundary, we note that quantum mechanics does not suffer from the same problem.
 The entropy of any pure state in quantum mechanics is zero, not minus infinity.
@@
 This is the inherent problem with classical mechanics: perfect isolation at all scales.
-In the correct circumstances, perfect isolation is an internally consistent assumption.
-In the correct circumstances, recursive subdivision into smaller parts is an internally consistent assumption.
+In the correct circumstances, perfect isolation is a reasonable assumption.
+In the correct circumstances, recursive subdivision into smaller parts is a reasonable assumption.
 It is the joint assumption that the system can be recursively subdivided into perfectly isolated parts that is physically untenable.\footnote{Note that the mathematical model is still perfectly consistent.
@@
 
-We have shown that classical mechanics lies on the implicit assumption of perfect isolation at all scales: every fragment of matter, every minuscule wave packet, can be produced, manipulated and measured with infinitesimal precision.
+We have shown that classical mechanics depends on the implicit assumption of perfect isolation at all scales: every fragment of matter, every minuscule wave packet, can be produced, manipulated and measured with infinitesimal precision.
 This is not only untenable, but it is also in conflict with thermodynamics.

### 2023-05-19 Articles/2023-PhysicalDifferentialTopology/PhysicalDifferentialTopology.tex
Minor fixes

@@
 Experimentally, we do not measure directly a mass density, but rather we measure mass and volume of a finite region and take the ratio.
-We then imagine to take smaller and smaller regions, and the mass density is ratio in the limit.\footnote{The Radon–Nikodym derivative, for example, follows this approach.} Ideally, our definitions should mirror this process to which there are at least to major advantages.
+We then imagine to take smaller and smaller regions, and the mass density is ratio in the limit.\footnote{The Radon–Nikodym derivative, for example, follows this approach.} This is similar for other quantities: we do not measure the magnetic field at a point but rather the magnetic flux over an ever smaller finite surface; we do not measure the force at a point but rather the work done by a force over an ever smaller finite path.
+Ideally, our definitions should mirror this process, which would have at least two major advantages.
 First, it is a lot easier to build physical intuition on the finite objects than on the local ones.
 Second, we can understand more clearly what physical requirements and idealizations are embedded in the definitions.
-For example, one basic assumption to define the mass density is that if we partition the volume, the total mass is the sum of the one found in each part.
+For example, one basic assumption to define the mass density is that if we partition the volume, the mass in the whole volume is the sum of the one found in each part.
 Given the equivalence of mass and energy, this is strictly true only if there is no interaction energy between the parts.
@@
 position and velocity) and others in derived units (e.g.
-velocity), therefore a change of definition in the first induces a change in the others.
-Since mathematics does not explicitly formalize units, it can do two things: either just define objects based on their transformation rules or look for objects that, by definition, follow the same transformation rules.\footnote{It seems that,in the mathematical world, the latter approach has been gaining favor with respect to the second.} To give an example of why this is problematic, in differential geometry one talks about tangent and cotangent bundle.
+velocity), therefore a change of definition in the first induces a change in the others.\footnote{As a more concrete example, consider the expression for work $W = \int_\lambda f_i dx^i$ as the line integral of the force.
+Work is an invariant meaning that its unit is independent of the the units used for the coordinates $x^i$.
+However, the units used for the components of the force will necessarily be the ration between the work unit and the corresponding unit for the coordinate.} Since mathematics does not explicitly formalize units, it can do two things: either just define objects based on their transformation rules or look for objects that, by definition, follow the same transformation rules.\footnote{It seems that,in the mathematical world, the latter approach has been gaining favor with respect to the second.} To give an example of why this is problematic, in differential geometry one talks about tangent and cotangent bundle.
 Roughly speaking, the first pairs a point with a vector and the second pairs a point with a one form.

### 2023-05-22 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
Update classical-phil-article-Erkenntnis.tex
Added in some references omitted for review. Running bibtex generates a `references.bib' file which is used to compile the bibliography.

@@
 Url = {https://doi.org/10.1111/phpr.12681}, 
+}
+
+@book{aop-book,
+	author    = {Gabriele Carcassi and Christine A. Aidala},
+	title     = {Assumptions of Physics},
+	publisher = {Michigan Publishing},
+	year      = 2021,
+	isbn      = {978-1-60785-706-7},
+	doi       = {10.3998/mpub.12204707},
+	url       = {https://doi.org/10.3998/mpub.12204707}
+}
+
+@ARTICLE {AoPPhy1,
+	author  = "Gabriele Carcassi and Christine A. Aidala and David J. Baker and Lydia Bieri",
+	title   = "From Physical Assumptions to Classical and Quantum Hamiltonian and Lagrangian Particle Mechanics",
+	journal = "Journal of Physics Communications",
+	year    = "2018",
+	volume  = "2",
+    number  = "4",
+    pages   = "045026",
+	url     = "https://doi.org/10.1088/2399-6528/aaba25"
+}
+
+@ARTICLE {shpmp,
+	author  = "Gabriele Carcassi and Christine A. Aidala",
+	title   = "Hamiltonian Mechanics is Conservation of Information Entropy",
+	journal = "Studies in History and Philosophy of Modern Physics",
+	year    = "2020",
+	volume  = "71",
+    pages   = "60--71",
+	url     = "https://doi.org/10.1016/j.shpsb.2020.04.004"
+}
+
+@thesis{Huntthesis,
+ Author  = {Hunt, Josh},
+  Title   = {Symmetry and Reformulation: {O}n Intellectual Progress in Science and Mathematics},
+type = {Ph.D. Thesis},
+  Year    = {2022},
+ School = {University of Michigan},
+}
+
+@Article{HuntPSA,
+  Author  = {Hunt, Josh},
+  Journal = {Philosophy of Science},
+  Title   = {Understanding and Equivalent Reformulations},
+  Year    = {2021},
+  Volume = {88},
+   Number = {5}, 
+Pages     = {810--823},
 }
@@
 However, he also shows that if we consider a more general formulation of Hamiltonian mechanics on a symplectic manifold (which does not require a cotangent bundle), then the two formulations are \textit{not} categorically equivalent \parencites*[1182-83]{Barrett2}.
-Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation {\color{green} and illuminates why it is more general.}} Whereas North's argument relies on a problematic criterion for comparing structure based on the size of symmetry groups \parencites[]{Swanson}, our argument avoids this pitfall.
+Our construction of Hamiltonian mechanics in Section~\ref{derivation} matches this more general symplectic manifold formulation and illuminates why it is more general.} Whereas North's argument relies on a problematic criterion for comparing structure based on the size of symmetry groups \parencites[]{Swanson}, our argument avoids this pitfall.
 We show that a classical system needs to satisfy strictly fewer physical conditions in order to be Hamiltonian.
@@
 Hamiltonian mechanics thereby captures a wider range of systems than Lagrangian mechanics.
-{\color{green} Unlike Lagrangian mechanics, Hamiltonian mechanics provides a unique solution to the equations of motion even outside the hyper-regular domain.
-Hence, Lagrangian mechanics is restricted to the hyper-regular regime.} The construction at the heart of our argument provides a novel interpretation of classical mechanical systems in terms of their invariant phase space density.
+Unlike Lagrangian mechanics, Hamiltonian mechanics provides a unique solution to the equations of motion even outside the hyper-regular domain.
+Hence, Lagrangian mechanics is restricted to the hyper-regular regime.
+The construction at the heart of our argument provides a novel interpretation of classical mechanical systems in terms of their invariant phase space density.
 
@@
 Section~\ref{derivation} describes in detail the derivation of Hamiltonian mechanics for systems that satisfy two physical conditions: infinitesimal reducibility and deterministic/reversible evolution.
-{\color{green} Our derivation shows that even non-hyper-regular Hamiltonian systems are physically interpretable.} We then show that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
+Our derivation shows that even non-hyper-regular Hamiltonian systems are physically interpretable.
+We then show that Lagrangian systems must satisfy an additional physical condition, which we call motion equivalence.
 This completes our primary argument for privileging Hamiltonian mechanics.
@@
 We provide a geometric interpretation of the action principle that applies even outside the hyper-regular domain.
-We end by rebutting objections that arise from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
+We end by rebutting objections that arise from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.\footnote{This paper forms part of the ongoing \textit{Assumptions of Physics} project \parencites[]{aop-book}.
+This project aims to both (i) isolate the physical assumptions required for different frameworks in mathematical physics and (ii) consider their philosophical implications.}
 
@@
 Of various notions of relative fundamentality, we focus on an entirely non-mysterious one.
-Given two formalisms that capture a class of systems, if one is logically more general, then it is more fundamental.
-This aspect of fundamentality is familiar from mathematics.
+Given two formalisms that capture a class of systems, if one is logically more general, then it is more fundamental.\footnote{Hunt \parencites*[Ch.
+5]{Huntthesis} provides an expressivist conception of fundamentality, based on an attitude of \textit{being for privileging}.
+In terms of this conception, adopting our specific criterion here expresses an attitude of being for privileging frameworks that are logically more general.
+Hunt \parencites*[816]{HuntPSA} provides a separate reason for privileging Hamiltonian mechanics: in systems with an ignorable coordinate, Hamiltonian mechanics makes manifest conservation of momentum.} This aspect of fundamentality is familiar from mathematics.
 A more fundamental framework encompasses a less fundamental one as a special case.
@@
 We do this in Section~\ref{derivation} by recovering the mathematical formalisms of both Hamiltonian and Lagrangian mechanics from three physical conditions.
-{\color{green}Our derivation illuminates the relationship between Hamiltonian and Lagrangian mechanics.
+Our derivation illuminates the relationship between Hamiltonian and Lagrangian mechanics.
 As North \parencites*[]{North} argues, there is a sense in which Hamiltonian mechanics is more general than Lagrangian mechanics.
@@
 Additionally, one might wonder whether Lagrangian mechanics could be meaningfully formulated in an analogously general manner.
-Our derivation helps answer these questions: the physical principles that undergird the symplectic manifold formulation of Hamiltonian mechanics are simply more general than the principles required for Lagrangian mechanics.}\footnote{{\color{green}We thank a referee for suggesting this helpful way of framing the significance of our derivation.}} In this section, we describe how our argument works in principle.
+Our derivation helps answer these questions: the physical principles that undergird the symplectic manifold formulation of Hamiltonian mechanics are simply more general than the principles required for Lagrangian mechanics.\footnote{We thank a referee for suggesting this helpful way of framing the significance of our derivation.} In this section, we describe how our argument works in principle.
 Section~\ref{fundamentality} demonstrates some philosophical advantages of our criterion for relative fundamentality.
@@
 we cannot specify the whole state by specifying the states of its subsystems).\footnote{The conditions of infinitesimal (ir)reducibility are similar to various notions of (non-)separability in the literature on physical holism \parencites[46, 124]{Healey}.
-We adopt our preferred terminology to avoid controversies surrounding how best to characterize separability.} We show in Section~\ref{infinitesimal} that {\color{green} if one chooses to model continuous quantities using a manifold, then this manifold must be symplectic.} To recover Hamiltonian mechanics, it then suffices to identify when trajectories on phase space obey Hamiltonian evolution (i.e.
+We adopt our preferred terminology to avoid controversies surrounding how best to characterize separability.} We show in Section~\ref{infinitesimal} that if one chooses to model continuous quantities using a manifold, then this manifold must be symplectic.
+To recover Hamiltonian mechanics, it then suffices to identify when trajectories on phase space obey Hamiltonian evolution (i.e.
 evolve according to a symplectomorphism).
@@
 We thereby characterize Hamiltonian mechanics as the mathematical framework for infinitesimally reducible systems with deterministic and reversible evolution.
-{\color{green} Importantly, our derivation encompasses Hamiltonian mechanics in the non-hyper-regular regime as well.}
+Importantly, our derivation encompasses Hamiltonian mechanics in the non-hyper-regular regime as well.
 
@@
 Hence, motion equivalence entails a correspondence between velocity and momentum, which formally leads to the framework of Lagrangian mechanics (Section~\ref{motion}).
-{\color{green} Since a system satisfies motion equivalence if and only if it is hyper-regular, this condition illuminates the physical significance of hyper-regularity.
-Moreover, since Lagrangians lack unique solutions to the equations of motion outside the hyper-regular regime, we take ourselves to have recovered the physically significant Lagrangian systems.}\footnote{{\color{green}See \textcites[119]{Gelfand} for discussion of the fact that Lagrangians yield unique solutions only in the hyper-regular domain.}} As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
+Since a system satisfies motion equivalence if and only if it is hyper-regular, this condition illuminates the physical significance of hyper-regularity.
+Moreover, since Lagrangians lack unique solutions to the equations of motion outside the hyper-regular regime, we take ourselves to have recovered the physically significant Lagrangian systems.\footnote{See \textcites[119]{Gelfand} for discussion of the fact that Lagrangians yield unique solutions only in the hyper-regular domain.} As a result, Lagrangian mechanics requires that a further physical condition is satisfied, and hence it is less general than Hamiltonian mechanics.
 
@@
 Compared with North's \parencites*[]{North} account of comparative structure, our criterion has two advantages.
-First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which scientific anti-realists, pragmatists, and even many naturalists view with suspicion.
+First, it avoids metaphysical disputes surrounding perfectly natural or joint-carving properties, which scientific anti-realists, pragmatists, and even some scientific realists view with suspicion.
 Secondly, it avoids a technical objection that Swanson and Halvorson \parencites*[]{Swanson} levy against North's more precise mathematical criterion for comparing structure.
@@
 Our derivation has three main steps.
-Section~\ref{infinitesimal} shows that {\color{green} if one represents continuous quantities using a manifold, then this manifold must be symplectic.} For systems with deterministic and reversible evolution, Section~\ref{deterministic} shows they must evolve according to a symplectomorphism.
+Section~\ref{infinitesimal} shows that if one represents continuous quantities using a manifold, then this manifold must be symplectic.
+For systems with deterministic and reversible evolution, Section~\ref{deterministic} shows they must evolve according to a symplectomorphism.
 Hence, we derive the formalism of Hamiltonian mechanics from the conditions of infinitesimal reducibility and deterministic/reversible evolution.
 Section~\ref{motion} shows how systems that satisfy the further condition of motion equivalence can be treated within the formalism of Lagrangian mechanics.
-{\color{green}In fact, when motion equivalence fails, Lagrangian mechanics does not yield unique solutions to the equations of motion.}
+In fact, when motion equivalence fails, Lagrangian mechanics does not yield unique solutions to the equations of motion.
 
@@
 First, we will show that a classical phase space arises if a system satisfies the following physical condition:
-\newpage % % %%remember to take out!
+%\newpage % % %%remember to take out!
 
@@
 This assumption characterizes classical systems as those whose infinitesimal parts completely specify the state of the system.
-We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.} {\color{green} Ultimately, this condition entails that we can describe the state of the whole system as a statistical distribution over the state space of its parts.
+We define a \textit{classical particle} as one of these infinitesimal parts, namely the limit of recursive subdivision.\footnote{This definition of a classical particle differs from viewing point-particles as extensionless \parencites[]{Butterfieldpoints} or zero-dimensional \parencites[]{Wilson}.} Ultimately, this condition entails that we can describe the state of the whole system as a statistical distribution over the state space of its parts.
 If we model the state space of the parts using a manifold, the distribution must be characterized by a density that depends on the state and not on the coordinates.
-From this condition, we show that the manifold must be symplectic.}
+From this condition, we show that the manifold must be symplectic.
+Readers satisfied with this overview can proceed to Section~\ref{deterministic}.
 % % Could add: readers who are satisfied with this overview can proceed to the next subsection
@@
 That is, the state of the whole system is a distribution over its infinitesimal parts.
-This distribution describes the proportion of classical particles in any particular state from $\mathcal{S}$.
+This distribution describes the proportion of classical particles in any particular state from $\mathcal{S}$.\footnote{For further details, see \textcites[]{AoPPhy1} and \textcites{shpmp}.}
 %ref 1: worth clarifying relationship between particle statespace S and system state space C.
@@
 
-When the system is characterized by continuous quantities, {\color{green}$\mathcal{S}$ can be taken to be a manifold}.\footnote{{\color{green}Note that the use of manifolds constitutes a choice of how to represent the physics.
+When the system is characterized by continuous quantities, $\mathcal{S}$ can be taken to be a manifold.\footnote{Note that the use of manifolds constitutes a choice of how to represent the physics.
 One could make another choice, for example, by representing the state variables not as local coordinates but as an algebra of operators that act on the state.
-In this case, we conjecture that a similar argument would demonstrate an analogue of a symplectic structure for this alternative mathematical representation.}} For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
+In this case, we conjecture that a similar argument would demonstrate an analogue of a symplectic structure for this alternative mathematical representation.} For $U \subseteq \mathcal{S}$, we define $\mu(U) := \int_U \rho_c(s) d\mathcal{S}$ to measure the number of particles in region $U$.
 \textit{State variables} $\xi^a$ are any set of quantities that fully identify states, so that $\xi$: $\mathcal{S} \to \mathbb{R}^n$ is an invertible function between each state and the numeric values of the variables.\footnote{In the language of manifolds, these $\xi^a$ are called ``coordinates'' and $\xi$ is a coordinate map.
@@
 \noindent In our case, this applies to both the state of the whole system and the state of its parts.
-{\color{green} Under this assumption, we show that the symplectic structure must be preserved over time, which gives us Hamiltonian evolution.}
+Under this assumption, we show that the symplectic structure must be preserved over time, which gives us Hamiltonian evolution.
 
@@
 In summary, we have shown that any system that satisfies the conditions of both infinitesimal reducibility and deterministic/reversible evolution is a classical Hamiltonian system.
-{\color{green} Note in particular that our derivation is not restricted to only hyper-regular Hamiltonians.
-As we will see, this is a key difference between Hamiltonian and Lagrangian mechanics.}
+Note in particular that our derivation is not restricted to only hyper-regular Hamiltonians.
+As we will see, this is a key difference between Hamiltonian and Lagrangian mechanics.
 
@@
 \noindent If we focus on the parts of the system, this means that for every evolution in phase space there should be one and only one trajectory in physical space.
-{\color{green}This assumption means that momentum can be expressed as a function of position and velocity, which mathematically corresponds to the case in which the Hamiltonian is hyper-regular.
+This assumption means that momentum can be expressed as a function of position and velocity, which mathematically corresponds to the case in which the Hamiltonian is hyper-regular.
 In this regime we find Lagrangian mechanics.
-Moreover, to be able to express the symplectic form (and therefore the density) over position and velocity, we find that momentum must be a linear function in velocity, which constrains the Hamiltonian to be that of a massive particle under scalar and vector potential forces.}
+Moreover, to be able to express the symplectic form (and therefore the density) over position and velocity, we find that momentum must be a linear function in velocity, which constrains the Hamiltonian to be that of a massive particle under scalar and vector potential forces.
 
@@
 Since the Jacobian is invertible, the Hessian must be non-zero everywhere, and therefore it must have the same sign everywhere---which we take to be positive.
-{\color{green} This, by definition, means that the Hamiltonian is hyper-regular if and only if weak equivalence is satisfied.} Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables.
+This, by definition, means that the Hamiltonian is hyper-regular if and only if weak equivalence is satisfied.
+Weak equivalence allows us to express particle states in terms of either position/momentum or position/velocity variables.
 These are exactly the cases where a Lagrangian can be constructed, and where the Lagrangian leads to a unique solution (via the stationary action principle).

### 2023-05-25 Articles/2020-ClassicalMechanicsPhil/Submissions/ERKE/classical-phil-article-Erkenntnis.tex
Update classical-phil-article-Erkenntnis.tex
Added Acknowledgments and moved mention of AoP project there. Deleted all comments, so as to send source file to the editorial office.

@@
 %!TEX encoding = UTF-8 Unicode
+
+%% To compile bibliography, run Bibtex, which will generate a file `references.bib' which is called in the References section. 
+%The file calls one diagram, labeled `Diagram.png'
 
@@
 \geometry{top=4cm, bottom=4cm, left=2.5cm, right=2.5cm}
-%\geometry{top=4cm, bottom=4cm, left=3cm, right=4cm} %close to actual philsci margins
-
-%\usepackage{endnotes}
-%\let\footnote=\endnote %Erkenntnis requires end-notes, but recent papers don't seem to follow this in the pdf!
- % %To call the endnotes, use command \theendnotes before calling reference page
 
@@
 \iffalse
-\titleformat{\section}[runin]% runin puts it in the same paragraph
-        {\normalsize\bfseries}% formatting commands to apply to the whole heading
+\titleformat{\subsubsection}[runin]% for acknowledgements 
+{\normalsize\bfseries}%
  {\thesection.}% the label and number
@@
 
-\titlespacing*{\section}{0pt}{2.5mm}{2.5mm}%spacing from left margin, spacing above header, spacing on right of header
 \fi
@@
 \fancyhead[R]{\thepage} %page number placement
-\fancyhead[C]{\textsc{Hamiltonian Privilege}}
-%\fancyhead[EC]{\textsc{AUTHOR NAMES}}%Even page header 
+%\fancyhead[C]{\textsc{Hamiltonian Privilege}}
+\fancyhead[OL]{\small{Hamiltonian Privilege}}
+\fancyhead[ER]{\small{Hunt, Carcassi, and Aidala}}%Even page header 
 \fancyfoot[L,R,C]{}
@@
   author  = {Butterfield, Jeremy N.},
-  journal = {British Journal for Philosophy of Science},
+  journal = {The British Journal for the Philosophy of Science},
   title   = {Against Pointillisme about Mechanics},
@@
 	year      = 2021,
-	isbn      = {978-1-60785-706-7},
 	doi       = {10.3998/mpub.12204707},
@@
 \usepackage[backend=bibtex, authordate, url=false, doi=false, eprint=false]{biblatex-chicago}
-%\addbibresource{bibliography.bib}
+
 \addbibresource{references.bib} %call the bib file generated above
@@
   \clearfield{number}%for convenience/consistency
+  \clearfield{isbn}%
 }
@@
 	pdftitle={Hamiltonian Privilege},
-	pdfauthor={},
+	pdfauthor={Hunt, Carcassi, and Aidala},
 	pdfsubject={},
@@
 %\author{\Large \textsuperscript{*\textdagger}}
-\date{\vspace{-7.9ex} \today}
+\date{\vspace{-7.9ex} Josh Hunt, Gabriele Carcassi, and Christine Aidala \\ \today}
 \maketitle
@@
 
-%\noindent\rule{\textwidth}{1.25pt}
 \bs
@@
 \es
-%\noindent\rule{\textwidth}{1.25pt}
-
-%\vspace{2mm}
 
@@
 We provide a geometric interpretation of the action principle that applies even outside the hyper-regular domain.
-We end by rebutting objections that arise from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.\footnote{This paper forms part of the ongoing \textit{Assumptions of Physics} project \parencites[]{aop-book}.
-This project aims to both (i) isolate the physical assumptions required for different frameworks in mathematical physics and (ii) consider their philosophical implications.}
+We end by rebutting objections that arise from Curiel's \parencites*[]{Curiel} privileging of Lagrangian mechanics.
 
@@
 Given two formalisms that capture a class of systems, if one is logically more general, then it is more fundamental.\footnote{Hunt \parencites*[Ch.
-5]{Huntthesis} provides an expressivist conception of fundamentality, based on an attitude of \textit{being for privileging}.
+6]{Huntthesis} provides an expressivist conception of fundamentality, based on an attitude of \textit{being for privileging}.
 In terms of this conception, adopting our specific criterion here expresses an attitude of being for privileging frameworks that are logically more general.
-Hunt \parencites*[816]{HuntPSA} provides a separate reason for privileging Hamiltonian mechanics: in systems with an ignorable coordinate, Hamiltonian mechanics makes manifest conservation of momentum.} This aspect of fundamentality is familiar from mathematics.
+Hunt \parencites*[816]{HuntPSA} considers a separate reason for privileging Hamiltonian mechanics: in systems with an ignorable coordinate, Hamiltonian mechanics makes manifest conservation of momentum.} This aspect of fundamentality is familiar from mathematics.
 A more fundamental framework encompasses a less fundamental one as a special case.
@@
 As Gilton notes, ``many philosophers of science have moved away from thinking of fundamentality in science in terms of the very small toward thinking of it in terms of comprehensiveness, or breadth of applicability'' \parencites*[637]{Gilton2021}.
-% % Removed: Similarly, an axiomatization that includes more axioms is less fundamental than one which drops these axioms while maintaining the same expressive power.
-%Worry: what if  both axiom systems cover the same class of systems? then seemingly no difference in generality? 
-%we must provide a criterion for one framework being more or less fundamental than another
-%\footnote{In Section~\ref{density}, we embrace a related non-metaphysical notion of fundamentality that pertains to physical entities rather than theoretical frameworks.}
 
@@
 Additionally, one might wonder whether Lagrangian mechanics could be meaningfully formulated in an analogously general manner.
-Our derivation helps answer these questions: the physical principles that undergird the symplectic manifold formulation of Hamiltonian mechanics are simply more general than the principles required for Lagrangian mechanics.\footnote{We thank a referee for suggesting this helpful way of framing the significance of our derivation.} In this section, we describe how our argument works in principle.
+Our derivation helps answer these questions: the physical principles that undergird the symplectic manifold formulation of Hamiltonian mechanics are simply more general than the principles required for Lagrangian mechanics.\footnote{We thank a reviewer for suggesting this helpful way of framing the significance of our derivation.} In this section, we describe how our argument works in principle.
 Section~\ref{fundamentality} demonstrates some philosophical advantages of our criterion for relative fundamentality.
-%We will, however in Section 5 give a criterion for an entity to be physically fundamental
-%These bolster our claim that this is an appropriate sense of fundamentality for philosophers to focus on.
-
-% From Healey 2007, (weak) separability says that ``any physical process occupying space-time region R supervenes upon an assignment of qualitative intrinsic physical properties at space-time points in R (and/or in arbitrarily small neighborhoods of those points)'' [46]
-% Non-separability: ``some physical process occupying a region R of space-time is not supervenient upon an assignment of qualitative intrinsic physical properties at space-time points in R'' [124]
-% Strong non-separability: ``some physical process occupying a region R of space-time is not supervenient upon an assignment of qualitative intrinsic physical properties at points of R and/or in arbitrarily small neighborhoods of those points'' [124].
 
@@
 Before presenting the details of our argument in Section~\ref{derivation}, we describe how we avoid problems that arise in North's privileging of Hamiltonian mechanics.
-%, although Curiel seems to view it as a virtue of the Lagrangian formulation that it handles dissipative systems with fewer modifications.
 
@@
 Secondly, it avoids a technical objection that Swanson and Halvorson \parencites*[]{Swanson} levy against North's more precise mathematical criterion for comparing structure.
-% Could cite Woodward, or Cohen/Callender for skepticism about perfectly natural properties. Includes naturalistically-minded philosophers
-%provided a theoretical framework captures the relevant behavior, it is more fundamental if it requires fewer conditions to hold.
 
@@
 Motion equivalence places an additional constraint on the Hamiltonian, a constraint that need not be satisfied in general.
-%Section~\ref{density} provides a further argument for viewing Hamiltonian mechanics as more fundamental.
-We argue that (i) invariant phase space densities are physically fundamental and (ii) Hamiltonian mechanics provides the proper setting for representing these densities.
-%is this the actual phase space density vs. density for all possible trajectories?
 
@@
 We thereby elucidate the physical significance of these formalisms.
-%details of this figure will become clear as we go along.
 
@@
 
-%Like many arguments in physics, our derivation falls short of some norms on mathematical rigor. 
-%optional to do: give a gloss on notion of regulative assumption (Kant and Peirce)
-
 \subsection{Infinitesimal reducibility}
@@
 First, we will show that a classical phase space arises if a system satisfies the following physical condition:
-%\newpage % % %%remember to take out!
 
@@
 Readers satisfied with this overview can proceed to Section~\ref{deterministic}.
-% % Could add: readers who are satisfied with this overview can proceed to the next subsection
-% % % For further discussion, see \textcites{shpmp}. [omit this for review].
 
@@
 This distribution describes the proportion of classical particles in any particular state from $\mathcal{S}$.\footnote{For further details, see \textcites[]{AoPPhy1} and \textcites{shpmp}.}
-%ref 1: worth clarifying relationship between particle statespace S and system state space C.
-Is there an S_i for each i-th particle?
-And explain why we are not focusing on the statespace of the many particle system, e.g.
-a cartesian product of state spaces for each particle
-
-%If $\mathcal{S}$ is a discrete space with $U \subseteq \mathcal{S}$, we define $\mu(U) = \sum\limits_{s \in U} \rho_c(s)$ so that $\mu(\mathcal{S})$ gives us the size of the system and $\mu(U)/\mu(\mathcal{S})$ gives us the fraction of the particles that can be found in the region $U$.\footnote{The discrete case is far less interesting because discrete quantities cannot change continuously over time, meaning that they must be constant.
-We discuss this elsewhere.}
-% % % For details see \textcites[]{AoPPhy1}. [omit this for review]}
-% % XXX could try to find justification/derivation for the state space being a manifold in the Kevin Kelly book, or could cite the math paper with Greenfield; proposition 6.2.
-Try to include a better gloss/motivation for this claim (basic idea: if time is continuous, trajectories in state space must be continuous, local structure must be compatible).
-Curiel paper might also have a brief justification/motivation, if I recall
 
@@
 It allows us to compare and track the amount of matter in different possible states.
-%The use of manifolds is, technically, a choice of how the physics is represented
 
@@
 Our next assumption involves the relationship between state variables and measurement units.
-%Since these transformation properties are required to use a density to model a classical mechanical system, it follows that having the structure of a symplectic manifold is likewise required.  
-%From the preceding two equations, we see that $\int_{\xi(U)} \rho_{c, \xi} (\xi^a) d \xi^a = \int_{\xi(U)} \rho_{c, \hat{\xi}} (\hat{\xi}^b) \left|\frac{\partial \hat{\xi}^b}{\partial \xi^a} \right|  d \xi^a$.
-% $d \xi^a$ and $d \hat{\xi}^b$ are the
 
@@
 Since the unit variables define the units for all state variables, this change must induce a unique change of state variables $\hat{\xi}^b = \hat{\xi}^b(\xi^a)$ for all variables.
-%to do: cite susan sterrett work on units as a `logical constraint'; , i.e. they fix the units for all state variables. 
-%ref 2: mention that any transformation on the unit variables `is required to extend to a unique transformation of all the coordinates'
 
@@
 Hence, we see that $\left|\frac{\partial \hat{k}^\beta}{\partial k^\alpha} \right| = \left|\frac{\partial q}{\partial \hat q} \right|$.
-%  $\left|\frac{\partial \hat{q}}{\partial q}\right|^{-1} = \left|\frac{\partial q}{\partial \hat q} \right|$.
 
@@
 
-%on needing a two-form to measure areas, could possibly cite Strang 1993, ch. 5 or Mac Lane & Birkhoff (1999, Theorem IX.2.2). 
-%note: an infinitesimal area counts the number of possible states in a region of phase space, as given by the two state variables that compose the degree of freedom
-
 Next, we show that $\omega$ must be closed, using the fact that the degrees of freedom are independent.
@@
 
-%Moreover, the units we choose for $\omega$ are the units we choose to count states over an independent degree of freedom. This agrees with the practice in statistical mechanics.
-
 We can thereby understand canonical coordinates as a set of state variables that express the density in the correct units.
@@
 Therefore symplectic structure is exactly the structure needed to define an invariant density.
-
-%Section~\ref{density} describes further interpretive consequences of this connection between invariant densities and symplectic structure.
 
@@
 In other words, the density remains the same throughout the evolution.
-%Let $\lambda: \mathbb{R} \to \mathcal{S}$ be the evolution over time of the state of a particle.
 
@@
 The evolution is therefore a symplectomorphism and hence corresponds to a Hamiltonian evolution.
-%Now consider the integral $\int_{\Sigma} \rho_c \omega(d\Sigma)$.
-Mapping the region $\Sigma$ to a new region, the fraction of the system found in both the old and new regions must be the same.
-Since both the integral and $\rho_c$ must remain constant during the evolution, the two-form $\omega$ must also remain the same.
 
@@
 As we will see, this is a key difference between Hamiltonian and Lagrangian mechanics.
-
-%Since, the only requirement that we have imposed on the Hamiltonian is that the equations yield a unique solution, our derivation applies to Hamiltonians that are not hyper-regular. }
-
-% since we have not imposed hyper-regularity,   }
-% [(ref 1): expand on how our argument relates to standard proofs of Liouville's theorem.
-Is it fair to say that we are privileging Hamiltonian mechanics because it forms the basis for a proof of Liouville's theorem?]
 
@@
 As we discuss in Section~\ref{Curiel}, Equation~\ref{weak_equivalence} corresponds to the kinematical constraints central to Curiel's \parencites*[]{Curiel} argument for privileging Lagrangian mechanics.}
-%; otherwise, the positions and velocities do not determine how the system evolves
 
@@
 We set:
-%\footnote{In Section~\ref{dependence}, we will see that $g^{ij}$ corresponds to the Riemannian metric, since $\partial v^i/\partial p_j$ is a function only of position.}
 \begin{equation}
@@
 Finally, we respond to Curiel's arguments against the fundamentality of Hamiltonian mechanics.
-% In the remainder of this essay, we provide further reasons for privileging Hamiltonian mechanics.
-The next section describes how Hamiltonian mechanics provides the natural setting for the invariant density, which characterizes how a classical object's parts are distributed.
-The invariant phase space density is more fundamental than the state variables, since conjugate momentum is a gauge-dependent quantity.
 
@@
 In short, we  argue that we have given a natural construction of Hamiltonian mechanics.
-%In short, we argue that we have given an eminently natural construction of Lagrangian mechanics, viewed as a special case of Hamiltonian mechanics.
-Moreover, our argument shows that Lagrangian mechanics is in some sense less natural than Hamiltonian mechanics because the former is contingent on having a bijection between velocity and conjugate momentum.
 
@@
 \label{action}
-
-% Somewhere below, emphasize that you have a unique trajectory only if the action is a minimum or maximum (if it is just stationary but not a minimum or maximum, then you do not have a unique trajectory)
 
@@
 
-% Helpful summary of Curiel's position: ``Hamiltonian mechanics does not respect the kinematical constraints intrinsic to abstract classical systems'' while ``Lagrangian mechanics always respect[s] the kinematical constraints of abstract classical systems'' \parencites*[304]{Curiel}.
-
 Curiel argues that Lagrangian mechanics represents classical mechanical systems in a \textit{natural} manner whereas Hamiltonian mechanics involves an essentially artificial construction \parencites*[270]{Curiel}.\footnote{Curiel characterizes ``natural'' in terms of natural coordinate systems, defined using the almost-tangent structure of the tangent bundle \parencites*[290-291]{Curiel}.
@@
 Thus, we take ourselves to have responded to Curiel's main criticism of Hamiltonian mechanics, namely that it must impose \textit{ad hoc} the condition (\ref{6.2}).
-% Note that our equation 2 (\ref{weak_equivalence}) is a special case (or less mathematically sophisticated formulation) of Curiel's kinematic conditions, i.e.
-his equation 4.3 in terms of the almost-tangent structure.
-Our equation 2b matches Curiel's equations 3.1a and 3.2b, in the case of free and non-free Newtonian particles, respectively.
 
@@
 Although Curiel argues this favors Lagrangian mechanics, it generates challenges for understanding non-hyper-regular Lagrangians.
-%\footnote{This challenge is particularly pressing for systems that are everywhere non-regular.
-In the case where a system is non-regular only at certain points, we can decompose the system into hyper-regular subsystems, each with its own Lagrangian (or Hamiltonian).}
-%\footnote{Barrett conjectures that outside the hyper-regular domain, Hamiltonian and Lagrangian formulations ``will be inequivalent according to any reasonable standard of equivalence.
-There is no natural way to translate a non-hyper-regular model of the one theory into a model of the other'' \parencites*[1179]{Barrett2}.}
 
@@
 Hence, our framework perspicuously distinguishes quantum from classical systems.
-We take these considerations to favor our preferred notion of a classical system.\footnote{Space precludes a detailed discussion of the quantum and statistical mechanical cases, which we hope to discuss elsewhere.
-Additionally, one can show that Hamiltonian mechanics provides the proper setting for the invariant density and a simple understanding of the gauge-dependency of conjugate momentum.}
-
-%\footnote{We provide details of the quantum and statistical mechanical cases elsewhere.
-Moreover, we show that Hamiltonian mechanics provides the proper setting for the invariant density and a simple understanding of the gauge-dependency of conjugate momentum.}
-%Preceding sections have illustrated many of these advantages.
-Hamiltonian mechanics provides the proper setting for the invariant density (Section~\ref{dependence}), a simple understanding of gauge transformations (Section~\ref{gauge_momentum}), and a general geometrical understanding of the principle of stationary action (Section~\ref{action}).
-% % %, see \textcites{AoPPhy1}. [omit for review].
+We take these considerations to favor our preferred notion of a classical system.\footnote{For more details on the quantum and statistical mechanical cases, see \textcites[]{AoPPhy1} and \textcites{shpmp}.
+Additionally, one can show that Hamiltonian mechanics provides both the proper setting for the invariant density and a simple understanding of the gauge-dependency of conjugate momentum.}
 
@@
 
-%\newpage
-
-%\theendnotes
-
-%\clearpage
+\subsubsection*{Acknowledgements}
+
+This paper forms part of the ongoing \textit{Assumptions of Physics} project \parencites[]{aop-book}.
+This project aims to both (i) isolate the physical assumptions required for different frameworks in mathematical physics and (ii) consider their philosophical implications.
+The authors would like to thank Gordon Belot for comments, along with two reviewers for this journal and four reviewers at two other journals, all of whom provided extremely helpful and constructive comments.
 
@@
 \defbibheading{myheading}[References]{%
-\centering \large \textbf{{#1}}%
+\flushleft \large \textbf{{#1}}
+%\centering \large \textbf{{#1}}%
 %\centering \footnotesize\textsc{{#1}}%
@@
 
-\printbibliography[heading=myheading]%see Stanley easier read, pages 63-64.
-
-%\printbibliography[title ={\centering \normalfont \textsc{References}}]
+\printbibliography[heading=myheading]
+
 \es

### 2023-05-25 Articles/2023-HilbertQuantum/HilbertQuantum.tex
HilbertQuantum - first draft

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amsfonts}

\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\def\>{\rangle}
\def\<{\langle}
\DeclareMathOperator{\erf}{erf}

\begin{document}

\title{The unphysicality of Hilbert spaces}
\author{TBD}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
TBD
\end{abstract}

\maketitle

\section{Introduction}

\section{On the physicality of Hilbert spaces}

First of all, let us clarify what exactly is the standard we should use to judge the physicality of a formal system. We say a mathematical definition is \textbf{physical} if it is grounded in some physical property or operation that is physically necessary. That is, we must be able to identify exactly what physical concept the mathematical definition is capturing, and show that the formal properties are indeed required by said physical concept. On the other hand, a mathematical definition is \textbf{unphysical} if it can be shown to require physical properties or operations that cannot be implemented physically.

For example, TODO example

To understand whether complex Hilbert spaces are suitable to represent physical entities, it is useful to break down the definition into three parts. Since a Hilbert space is a complete inner product space, we will look at the vector space structure, the inner product and the completeness separately.

\subsection{Vector space}

A vector space is a set where the elements can be summed and multiplied by a constant. In quantum mechanics, the set represents pure states and the linear combination represent superposition of states. One may already accept this as physical, as superpositions can be created in a lab when, for example, a particle goes through a beam splitter. However, there are two difficulties.

First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs. This is an area where different interpretations may differ, and we do not want our physical justification to depend on our interpretation. We want to decide whether the mathematical representation is physical regardless of the interpretation. Second, the actual states are not represented by vectors, but by rays. The superposition, then, is in terms of objects that are not themselves strictly physical. For these two reasons, we want to find another way an equivalent characterization of the linearity of the space that is fully physical.

Let us consider the space of density operators, which can be used to characterize both pure and mixed state. These objects are invariant under change of phase and are in one-to-one correspondence with the rays of the space, not with the vectors. The density operators have a linear structure that represent statistical mixtures. This is a necessary structure because preparing statistical mixtures is something we can always do physically, and therefore something a physically meaningful mathematical definition must provide. However, the linear structure of density operators is not the same as the linear structure of the Hilbert space.

When we write
$$c_1|\psi_1\> + c_2 |\psi_2\>$$
and 
$$p_1|\psi_1\>\<\psi_1| + p_2|\psi_2\>\<\psi_2|$$
the meaning of the $+$ symbols is different. The first operations is a complex linear combination of two pure states that returns another pure state, while the second is a real linear combination of two pure states that returns a mixed state. However, the two are so intertwined that one can be recovered by the other.

As we show in the appendix, a pure state $\psi$ is a superposition of two other pure states $\phi_1$ and $\phi_2$ if and only if there exists another pure states $\bar{\psi}$ such that we can mix $\psi$ and $\bar{\psi}$ to obtain a mixed state that can also be obtained by mixing $\phi_1$ and $\phi_2$. Geometrically, consider a Bloch sphere. If $\psi$ is a superposition of $\phi_1$ and $\phi_2$ then all three states are on the surface of the same sphere. The chord connecting $\phi_1$ and $\phi_2$ represents the space of all possible mixtures of $\phi_1$ and $\phi_2$. Now, take a point $\rho$ on that chord, and consider the line that passes through $\rho$ and $\psi$. The same line will traverse the sphere at another point, which we call $\bar{\psi}$. Since $\rho$ is on the chord that connects $\psi$ and  $\bar{\psi}$, it is a mixture of the two. Therefore the vector space structure can be recovered by knowing how pure state mix.

Note that the fact that the same statistical mixture can be decomposes in multiple ways is a characteristic feature of quantum mechanics, that is not present in classical mechanics. Using the same characterization, then, we consistently conclude that in classical mechanics there are no superpositions precisely because, given a pure state $\psi$, we can never find a corresponding state $\bar{\psi}$.

The argument, then, goes like this. On physical grounds, we must be able to construct mixture. On physical grounds, we must be able to tell whether the same mixture can be prepared in different ways. The vector space structure captures exactly which pairs of pure states produce indistinguishable mixtures. The vector space structure exactly captures that structure and therefore it is fully physically justified. \textbf{The vector space structure is physical.}

\subsection{Inner product}

The inner product defines a geometrical structure on top of the vector space. Namely, it gives an idea of size of each vector and of angle between vectors. In quantum mechanics, the square of the inner product
\begin{equation}
	p(\psi_1, \psi_2) = |\<\psi_1| \psi_2\>|^2.
\end{equation}
represent the probability of measuring one state having prepared another. One may already accept this as physical, as these probabilities define all observable features of quantum mechanics. There is, again, a difficulty.

Understanding the inner product as giving the probability of measurements also brings in one of the most controversial aspects of the theory: measurements. This is, again, a problem since we want our reasoning to be independent of interpretation. Another issue is that, technically, the inner product is linked to decomposition in linear combinations, to superposition, which are done at equal time. The probability above, instead, is a probability of transition.

Given that we have characterized superposition in terms of mixtures, we should see whether the inner product characterizes something fundamental about them, and indeed it does. Given two pure states, the entropy of the mixture is a strictly monotonic, and therefore invertible, function of the square of the inner product. The greatest entropy is given by mixing orthogonal states, which gradually decreases as the angle in the inner product space decreases. We can then understand the geometrical structure of the pure quantum states as an entropic structure, much like it happens for classical mechanics.\footnote{In classical mechanics, the Liouville measure returns the count of states within a region, which is used in the fundamental postulate of statistical mechanics. The symplectic form gives us the Liouville measure for each independent degree of freedom.}

The argument, then, goes like this. On physical grounds, for each statistical mixture, the thermodynamic entropy must be defined. Therefore it must be defined for equal mixtures of two pure states as well. The inner product exactly captures how that quantity combines and therefore it is fully physically justified. \textbf{The inner product structure is physical.}

\subsection{Completeness}

Completeness of a Hilbert space tells us that the space includes the limit of every Cauchy sequence. This property is more difficult to characterize physically. We can look at an equivalent characterization:\footnote{Roman 2008, p. 330 Theorem 13.8

	Roman, Stephen (2008), Advanced Linear Algebra, Graduate Texts in Mathematics (Third ed.), Springer, ISBN 978-0-387-72828-5} the space is complete if, given a series of vectors
$$ \sum _{k=0}^{\infty }u_{k}$$
such that
$$ \sum _{k=0}^{\infty }\|u_{k}\|<\infty \,$$
the series converges in the space. This is now in terms of linear combinations, and therefore it is essentially telling us that we can perform a linear superposition if the sum of the norm of the coefficient converge. We are basically extending superposition to infinite sums. Is this a physical requirement of just mathematical convenience?

As we saw before, superposition experimentally means splitting the state along multiple paths. The above mathematical expression means splitting the original state into infinitely many progressively small component, and be able to place them where we please. It also assumes, for example, that we can place them over an infinite spatial range. Since the limit must be in the space, we are literally saying that we can prepare states that are spread over infinite regions. This does not sound physical.

Again, given the difficulties of understanding superposition, let us reframe the result in terms of statistical quantities. Suppose that we are able to prepare pure states with arbitrary finite expectation value for position. Then completeness means we are also able to prepare pure states that falls off at infinity as $\frac{1}{x^2}$. The expectation value of $x$ would the integral of a function that falls off as $\frac{1}{x}$ which diverges. Clearly, this does not make physical sense. When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that observed value can be arbitrarily large, not that we can literally spread values over an infinite range.

The argument, then, goes like this. On physical grounds, given a quantity that can take arbitrarily large values, like position, for each possible value there must exist some state whose expectation matches that value. On mathematical grounds, we require completeness, which forces us to posit states with infinite expectation value. But this, on physical grounds, is not teneable. Therefore \textbf{completeness in unphysical.}

\subsection{Impact of unphysicality}

We are forced to conclude that Hilbert spaces have unphysical requirements when applied to quantum mechanics. We clearly see that the issue is the closure of operation over infinite sequences, which is not surprising.\footnote{Large branches of mathematics were developed to handle infinities, both large and small. Weyl stated ``Mathematics is the science of the infinite, its goal the symbolic comprehension of the infinite with human, that is finite, means.'' Hermann Weyl, "The Open World: Three Lectures on the Metaphysical Implications of Science," (1932) as quoted in Mind and Nature: Selected Writings on Philosophy, Mathematics, and Physics (2009) ed. Peter Pesic.} Now we have to understand how big of a problem it is.

On physical grounds, the state space of quantum system is the projective space of a complex inner product space. If the space is finite dimensional, the space is automatically complete.\footnote{This is exactly why, mathematically, completeness is nice as it makes infinite dimensional space behave like finite dimensional ones.} If not, we can always complete it by enlarging the space. In general, then, it is true that a quantum state is a ray of a Hilbert space, but the converse is not true: not all rays of the same Hilbert space will be valid quantum states.

This may not seem like a problem: we simply have a bigger set of object of what we strictly need. Unfortunately, in mathematics working on a bigger space is not always a good thing. For example, take the problem of defining volumes in three dimensional euclidean space $\mathbb{R}^3$. Formally, we want to find a measure $\mu(U)$ that returns the size of the region $U$. Naively, we may want to define it on the set of all possible subsets. In this case, Banach–Tarski paradox tells us that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls of the same size. Clearly, this is a problem. However, if we restrict ourselves with a measure on the Borel sets instead of all sets, we get a consistent definition. Therefore, in mathematics, bigger is not always better.\footnote{In another work, we show that the Borel sets are exactly those sets that can be associated with experimental procedures. In many cases, we have found that physical requirements naturally map to well behaved spaces.}

The above example is just a cautionary tale and it doesn't tell us what happens in this specific case. Therefore, let us give an example of the things that can go wrong. Unitary transformation play an important role in the theory of Hilbert spaces for at least two reasons. The first is that two Hilbert spaces that differ by a unitary transformation are mathematically equivalent. A change of representation, like a change of basis, is a unitary transformation. The second reason is that time evolution for an isolated system is represented by a unitary transformation. Let us see, then, the interplay between unitary transformations and our previous findings.

In a Hilbert space, we can always find a unitary transformation that maps one vector to any other. Therefore, there will be unitary transformations that map states with finite expectation values to infinite ones. This means that whether something has a finite expectation value or not will depend on the choice of representation. As a concrete example, consider the following two wave functions
\begin{align}
	\psi(x) &= \sqrt{\frac{e^{-x^2}}{\sqrt{\pi}}} \\
	\phi(x) &= \frac{1}{\sqrt{\pi(x^2 + 1)}}.
\end{align}
Both states are symmetric, and therefore will have zero expectation value for $x$. The first distribution is a Gaussian, and therefore all moments of the distribution, the expectation for $x^n$, are finite. The second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all. We can now find a change of variable $y(x)$ that transforms one distribution into the other. We need to set
\begin{equation}
	\begin{aligned}
	\int_{0}^{x} \psi^\dagger(x) \psi(x) dx &= \int_{0}^{y(x)} \phi^\dagger(x) \phi(x) dx \\
	\int_{0}^{x} \frac{e^{-x^2}}{\sqrt{\pi}} dx &= \int_{0}^{y(x)} \frac{1}{\pi(y^2 + 1)} dx \\
	\frac{\erf(x)}{2} &= \frac{\tan^{-1}(y)}{\pi} \\
	y &= \tan \left(\frac{\pi}{2}\erf(x)\right). \\
	\end{aligned}
\end{equation}
Therefore, thorough a variable change, we can change a state that has finite expectation value for all moments greater than the first into a state that has infinite expectation value for the same moments. Variable changes, in terms of the Hilbert space, are unitary transformation. Therefore the theory of Hilbert spaces, as applied to quantum mechanics, would tell us that the two representations are equivalent.

Experimentally, these are not equivalent. Identifying a state through expectation value is a common technique, called quantum tomography. If a unitary transformation can change finite expectations to infinite ones, we are effectively changing what can be distinguished through a tomography. Not only we cannot measure infinite values, but different finite values would get lumped into the same infinite value. In the above example, if we increase the variance, all even moments would change, yet, after the transformation, they would all be infinite. Therefore the idea that such unitary transformation links to physically equivalent representations is, on physics ground, untenable.

Moreover, we can create the following family of transformations
\begin{equation}
	y = \cos(\omega t) x + \sin(\omega t) \tan \left(\frac{\pi}{2}\erf(x)\right).
\end{equation}
This is clearly continuous in $t$ and, for $t=0$, we have the identity transformation and therefore, mathematically, there is no reason why this wouldn't be a valid time evolution. The effect is simply a stretching back and forth the wavefunctions, such that the state keeps oscillating between the above two functions. On physics grounds, however, we are going from finite to infinite expectations in finite time. This, again, is physically untenable.

We hope that this gives an idea of the type of issues embedded in the mathematical definitions. When performing actual calculations, we do not see these issues because the physics that we are describing makes us work in a corner of the structure where things are reasonable. But when one tries to prove theorems or general results on quantum theory, one currently is forced to consider the entire Hilbert space. It is not inconceivable to expect that results that would be reasonable physically fail to valid when enlarging the space of possible states to those that have infinite expectation values and the space of physical transformations to those that map finite to infinite expectations. We therefore wonder whether some of the issues we have in quantum field theory (e.g. the inability to find a valid measure for path integrals, inequivalent representation of interacting theories (?), infinities) are simply a consequence of the use, at a fundamental level, of unphysical objects.

\section{On the physicality of Schwartz spaces}

Suppose the reader agrees with our finding, the subsequent question is: what is the alternative? If we are only interested in observable with finitely many possible outcomes, we can use finite dimensional spaces. But as soon we have discrete quantities with arbitrarily large values, such as the number of particles, or quantities that are continuous, like position and momentum, we need infinite dimensional spaces. If we allow infinite dimensional spaces, we need some way of controlling what infinite linear combinations are allowed or not allowed. Cauchy completeness is one such way, but if that is not physical, what should we use?

We saw that the main problem is having expectation values that diverge, therefore the naive idea would be to require that the expectation value for all possible variables converge. This would require the expectation for $e^x$, $e^{e^x}$, $e^{e^{e^x}}$ and so on to be finite as well, which seems rather extreme. But what is a reasonable point to stop? If the idea is that we need to be able to reconstruct the state from expectation values, we should note that bounded distributions can be recovered from all the moments of the distribution, from the expectation of $x^n$ for every finite $n$. It seems reasonable, then, to require the expectation of all polynomial of position and momentum to be finite.

In quantum mechanics, this means requiring all expectations of $X^nP^m = -\imath \hbar x^n\partial_x ^m$ to be finite. This restricts our space to those functions that are infinitely smooth (i.e. all derivative exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions. Requiring finite expectation for all polynomials of position and momentum, then, does not give us the usual $L^2$ Hilbert space, it gives us the Schwartz space. Is this a better candidate?

First of all, the Schwartz space is a subspace of the $L^2$ Hilbert space taken as an inner product space. That is, all Schwartz functions are elements of the Hilbert space, they are closed under linear combinations and have the same inner product. It is not, however, complete, which is what we want. Moreover, it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with an arbitrary level of precision. Or, equivalently, that once a map is defined over the Schwartz space, the map is uniquely extended over the Hilbert space. This is a good hint that we have thrown away only unphysical elements.

If we restrict ourselves to the Schwartz space, then, the only unitary transformations we are interested in are those that map Schwartz function to Schwartz functions. The change of coordinate we defined in the previous section, then, is ruled out. Limits in terms of Cauchy sequences are not guaranteed, but limits in terms of expectation values are guaranteed: if we have a sequence of states such that the limit of the expectation converge, then we are guaranteed that the limit is also a Schwartz function.

The Schwartz space has also a very important property: it is the only space that is closed under the Fourier transform. That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property. This is not, in retrospect, a nice property that would be convenient for physicists. The theory of tempered distribution, which allows us to define the Dirac $\delta$-function, is mathematically based, at its core, over the Schwartz space. Therefore, this space already plays a fundamental role in the definitions of the tools we already use.

On the downside, many of the tools we use indeed assume a Hilbert space and many theorems would not hold without Cauchy completeness. There are two responses to this issue. First of all, there is no reason to prohibit that calculation are performed on the larger, unphysical space. When solving an integral, we do sometimes extend the domain for real to complex values, which are unphysical, because complex analysis has nicer mathematical features. The result so found, however, is still the valid result on the real domain. Similarly, we can pose the problem on the Schwartz space, extend to the Hilbert space for calculation, and then bring the result back to the Schwart space. However, there is a second way to look at this issue. Given that the focus has been squarely on Hilbert spaces, we do not know which mathematical results will hold, or even whether other results may be found. This is not something that the authors of this paper can answer and neither the mathematicians that were asked. This requires a serious research effort on the mathematical underpinning of multiple areas, which brings us to our closing point.

This effort is never going to happen if the physics community, as whole, do not generate enough interest on the subject. There may be a better series of mathematical tools and results that stem from a more precise specification of our physical requirements in terms of mathematical definitions. It is unlikely that the mathematical world will give us these tools unprompted. It is the job of us physicists to understand exactly what are ramifications of the definitions in terms of what physical objects would be left out or what unphysical objects would be brought in.

\section{Conclusion}

\bibliography{bibliography}

\section*{Appendix: math}

Let $\mathcal{H}$ be a complex vector space. Let $\psi, \phi_1, \phi_2 \in \mathcal{H}$ be pure states. Then $\psi$ is a superposition of $\phi_1$ and $\phi_2$, that is $\psi = c_1 \phi_1 + c_2 \phi_2$, if and only if there is another state $\bar{\psi}$ such that an equal mixture of $\psi$ and $\bar{\psi}$ is also a mixture of $\phi_1$ and $\phi_2$. That is
\begin{equation}
	\frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\bar{\psi}\>\<\bar{\psi}| = p_1|\phi_1\>\<\phi_1| + p_2|\phi_2\>\<\phi_2|
\end{equation}

Let $\mathcal{H}$ be a complex Hilbert space. Let $I : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the von Neumann entropy of the equal mixture of two pure states. That is
\begin{equation}
	I(\psi_1, \psi_2) = S\left(\frac{1}{2}|\psi_1\>\<\psi_1| + \frac{1}{2}|\psi_2\>\<\psi_2|\right).
\end{equation}
Let $p :  \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the square of the inner product. That is
\begin{equation}
	p(\psi_1, \psi_2) = |\<\psi_1| \psi_2\>|^2.
\end{equation}
Then $p$ and $I$ can be reconstructed from other. That is, there exists an invertible function $f$ such that
\begin{equation}
	I(\psi_1, \psi_2) = f(p(\psi_1, \psi_2)).
\end{equation}

\end{document}

### 2023-06-07 Articles/2023-HilbertQuantum/HilbertQuantum.tex
HilbertQuantum.tex
New intro.

@@
 \section{Introduction}
+
+Von Neumann's \textit{Mathematische Grundlagen der Quantenmechanik} (\citeyear{von_neumann_mathematische_1996}) crowned Hilbert spaces as (i) the unifying framework for the states in Schrödinger's wave mechanics and Heisenberg-Born-Jordan's matrix mechanics that is (ii) mathematically rigorous.
+Arthur Wightman, one of the people who followed von Neumann's aims to unify and axiomatize quantum field theory (\citeyear{wightman_quantum_1956}), took his book to be the fulfillment of a project Hilbert himself kicked off with his problems on the foundations of mathematics (\citeyear[p.
+157]{browder_hilberts_1976}): to axiomatize the physical sciences.
+Despite the undeniable importance of wanting to set physics on a firm mathematical footing, we must remember that by ``axioms,'' they meant ``physically motivated constraints'' more than statements that can be immune to question, rethink, and, if necessary, revise.\footnote{For a more extended discussion of von Neumann's axiomatics and their physical inputs, see \citep{redei_soft_2006}.}\\
+
+In this letter, we focus on an assumption that poses an array of interesting questions: that quantum states are vectors in Hilbert spaces.
+At first glance, some of their shortcomings (states are rays, not vectors; the Dirac delta is a distribution, not a function; some elements from $L^2(\mathbb{R})$ don't seem apt candidates to be wave functions) are well-known, but deemed unproblematic.
+Additionally, a handful of theorems (Gleason, Riesz, Wigner, Stone, Stone-von Neumann, Wigner, etc.) rely heavily on the structure of a Hilbert space.
+While these results are rarely mentioned when students are first introduced to quantum mechanics, they are crucial to substantiate the claim that Hilbert spaces are the right structure to characterize quantum states.
+Therefore, it's not a surprise that Hilbert spaces have become fossilized as parts of axiomatic systems\footnote{Algebraic quantum field theory (see, for instance, \citep{haag_local_1996}) inverts the narrative of ``quantum states are normed vectors in a Hilbert space, and observables are bounded, self-adjoint operators acting on that space'' to ``canonical (anti)commutation relations of observables generate an algebra, and states are assignments of expectation values to elements of the algebra.'' As such, it doesn't \textit{need} to talk about Hilbert spaces to do quantum mechanics, but it has resources, like the GNS construction, that allow their advocates to say that, since they're recovering Hilbert spaces from algebras of observables, what they're doing ends up being similar to the quantum mechanics that's more familiar to physicists.} without even considering whether we can do quantum mechanics with weaker assumptions.
+So, despite all their success, perhaps it's time to ask if using Hilbert spaces to describe quantum states, an assumption whose physical motivation was meant to be revised as time passed, has become a dogma.
 

### 2023-06-20 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Started to rewrite (and adding proof)

@@
 
-\usepackage{amsmath}
-\usepackage{mathrsfs}
-\usepackage{amsfonts}
+\usepackage{assumptionsofphysics}
 
-\usepackage{graphicx}
-\usepackage{hyperref}
-\hypersetup{
-	colorlinks=true,
-	citecolor=blue,
-	urlcolor=blue,
-	linkcolor=blue
-}
-\urlstyle{same}
-\frenchspacing
-
+%\usepackage{amsmath}
+%\usepackage{mathrsfs}
+%\usepackage{amsfonts}
+%
+%\usepackage{graphicx}
+%\usepackage{hyperref}
+%\hypersetup{
+%	colorlinks=true,
+%	citecolor=blue,
+%	urlcolor=blue,
+%	linkcolor=blue
+%}
+%\urlstyle{same}
+%\frenchspacing
+%
 \def\>{\rangle}
@@
 
-First of all, let us clarify what exactly is the standard we should use to judge the physicality of a formal system.
-We say a mathematical definition is \textbf{physical} if it is grounded in some physical property or operation that is physically necessary.
-That is, we must be able to identify exactly what physical concept the mathematical definition is capturing, and show that the formal properties are indeed required by said physical concept.
+Let us first clarify the standard we believe should be usde to judge the physicality of a formal definition.
+Suppose we have a physical system under study in a given condition.
+We say a mathematical definition is \textbf{physical} if it is the proper mathematical characterization of the physical system.
+That is, if it can be shown that the mathematical definition is needed to capture and only capture a particular aspect of the physical system.
 On the other hand, a mathematical definition is \textbf{unphysical} if it can be shown to require physical properties or operations that cannot be implemented physically.
 
-For example, TODO example
+For example, suppose we have a set of physically distinguishable objects, meaning that we can tell any two apart through an experimental test.
+Experimentally, the best we can do is collect finite information in finite time.
+If we are given an arbitrarily long time, we can collect an arbitrary amount of finite information.
+An element of the set, then, must be identified by at most a countably infinite sequence of ones and zeros, so that, at some point, we would be able to tell the difference between the two sequences and therefore the objects.
+The set of countable sequences of zeros and one can be understood as the binary expansion of real numbers between zero and one, which has cardinality of the continuum.
+Therefore, any definition that would require a set of physical objects to have cardinality greater than the continuum, then, is \textbf{unphysical}.\footnote{For example, non separable Hilbert spaces have cardinality greater than the continuum.}
 
-To understand whether complex Hilbert spaces are suitable to represent physical entities, it is useful to break down the definition into three parts.
-Since a Hilbert space is a complete inner product space, we will look at the vector space structure, the inner product and the completeness separately.
+Naturally, this does not mean that all sets with cardinality of the continuum or lower are physical.
+In general, testing physicality is harder as we need to check that all the mathematical features map to physical feature and no physical feature is left behind.
+Clearly, postulating that ``the state space of a quantum states has cardinality no grater than that of the continuum'' is certainly \textbf{physical}.
+We now need to check the rest.
 
-\subsection{Vector space}
+\subsection{Inner vector space}
 
-A vector space is a set where the elements can be summed and multiplied by a constant.
-In quantum mechanics, the set represents pure states and the linear combination represent superposition of states.
-One may already accept this as physical, as superpositions can be created in a lab when, for example, a particle goes through a beam splitter.
-However, there are two difficulties.
+Recall that a Hilbert space is a complete inner product space.
+That is, it is a vector product, with an inner product, closed under Cauchy sequences.
+Let us look at the first two structure briefly, as they are straight forward.
+We will leave more detail in the appendix.
 
-First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs.
-This is an area where different interpretations may differ, and we do not want our physical justification to depend on our interpretation.
-We want to decide whether the mathematical representation is physical regardless of the interpretation.
-Second, the actual states are not represented by vectors, but by rays.
-The superposition, then, is in terms of objects that are not themselves strictly physical.
-For these two reasons, we want to find another way an equivalent characterization of the linearity of the space that is fully physical.
+Roughly speaking, a vector space is a set whose elements can be combined linearly: can be multiplied by constants and summed.
+If one regards superpositions as physical requirements of quantum theory, one may think that this structure is physical.
+However, the ability to interpret linear combination as superpositions requires the ability to keep track of normalization.
+If we can only know whether two vectors are in the same ray, without knowing which ones are normalized, we are not able to recover the idea of a probability amplitude, as those are the coeffcients of a linear combination between normalized states.
+Therefore the vector space structure, by itself, is not physical.
 
-Let us consider the space of density operators, which can be used to characterize both pure and mixed state.
-These objects are invariant under change of phase and are in one-to-one correspondence with the rays of the space, not with the vectors.
-The density operators have a linear structure that represent statistical mixtures.
-This is a necessary structure because preparing statistical mixtures is something we can always do physically, and therefore something a physically meaningful mathematical definition must provide.
-However, the linear structure of density operators is not the same as the linear structure of the Hilbert space.
+If we add the inner product, we can define the norm of vectors and therefore we can indeed express superpositions.
+The inner product structure also defines probabilities of transitions.
+If one is comfortable in considering superpositions and the Born rule to be physical requirements of quantum mechanics, then the inner product structure is \textbf{physical}.
+However, some may not consider superpositions as actual physical, but more as mathematical constructs; others may object that the Born rule, in its standard interpretation, defines probability of transitions while a state space should be defined at equal time; others can argue that the states are defined by the rays, and therefore the requirements should only be in terms of those objects.
+For those, we have an alternative, more necessary justification, which we include in the appendix, and can be summarized as follows.
 
-When we write
-$$c_1|\psi_1\> + c_2 |\psi_2\>$$
-and 
-$$p_1|\psi_1\>\<\psi_1| + p_2|\psi_2\>\<\psi_2|$$
-the meaning of the $+$ symbols is different.
-The first operations is a complex linear combination of two pure states that returns another pure state, while the second is a real linear combination of two pure states that returns a mixed state.
-However, the two are so intertwined that one can be recovered by the other.
+Given that, experimentally, no preparation or measurement is perfect, a physical theory must define the space of statistical ensembles.
+Mathematically, these form a convex set, that physically represents a set that allows statistical mixtures.
+The geometry of the space of statistical ensembles is uniquely defined by the inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtues.\footnote{There is a rich literature on Generalized Probability Theory in which people reconstruct Quantum Theory by imposing requirements on convex spaces.} For example, we can characterize superposition in the following way: a pure state $\psi$ is a superposition of two other pure states $\phi_1$ and $\phi_2$ if and only if there exists another pure states $\bar{\psi}$ such that we can mix $\psi$ and $\bar{\psi}$ to obtain a mixed state that can also be obtained by mixing $\phi_1$ and $\phi_2$.
+That is, the ability to create superposition is conceptually equivalent to the ability to prepare the same mixture in different ways.
+The inner product can be shown to be equivalent to determining the entropy of all equal mixtures of state pairs.
 
-As we show in the appendix, a pure state $\psi$ is a superposition of two other pure states $\phi_1$ and $\phi_2$ if and only if there exists another pure states $\bar{\psi}$ such that we can mix $\psi$ and $\bar{\psi}$ to obtain a mixed state that can also be obtained by mixing $\phi_1$ and $\phi_2$.
+If we require, then, that the theory reproduces the properties of quantum mixtures, how they can be decomposed in terms of pure states and what their entropy is, the inner product space is recovered.
+As we said, a physical theory must tell us how to properly characterized statistical mixture, therefore the requirement is necessitated by the physics.
+\textbf{The inner product space structure is physical}.
+
+----
+
  Geometrically, consider a Bloch sphere.
@@
 
+% TODO: hint that the unitary transformations on Schwartz space are a subset of the ones on the Hilbert, and therefore they have unique inverse; operators we are interested in are uniquely extended to the Hilbert space.
+
 \bibliography{bibliography}
 
-\section*{Appendix: math}
+\newcommand{\pj}[1] {\underbar{$#1$}}
 
+\section*{Appendix}
+
+We give the full mathematical detail in support of the argument laid out in the main paper
+
+A vector space is a set where the elements can be summed and multiplied by a constant.
+In quantum mechanics, the set represents pure states and the linear combination represent superposition of states.
+One may already accept this as physical, as superpositions can be created in a lab when, for example, a particle goes through a beam splitter.
+However, there are two difficulties.
+
+First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs.
+This is an area where different interpretations may differ, and we do not want our physical justification to depend on our interpretation.
+We want to decide whether the mathematical representation is physical regardless of the interpretation.
+Second, the actual states are not represented by vectors, but by rays.
+The superposition, then, is in terms of objects that are not themselves strictly physical.
+For these two reasons, we want to find another way an equivalent characterization of the linearity of the space that is fully physical.
+
+Let us consider the space of density operators, which can be used to characterize both pure and mixed state.
+These objects are invariant under change of phase and are in one-to-one correspondence with the rays of the space, not with the vectors.
+The density operators have a linear structure that represent statistical mixtures.
+This is a necessary structure because preparing statistical mixtures is something we can always do physically, and therefore something a physically meaningful mathematical definition must provide.
+However, the linear structure of density operators is not the same as the linear structure of the Hilbert space.
+
+When we write
+$$c_1|\psi_1\> + c_2 |\psi_2\>$$
+and 
+$$p_1|\psi_1\>\<\psi_1| + p_2|\psi_2\>\<\psi_2|$$
+the meaning of the $+$ symbols is different.
+The first operations is a complex linear combination of two pure states that returns another pure state, while the second is a real linear combination of two pure states that returns a mixed state.
+However, the two are so intertwined that one can be recovered by the other.
+
+As we show in the appendix, a pure state $\psi$ is a superposition of two other pure states $\phi_1$ and $\phi_2$ if and only if there exists another pure states $\bar{\psi}$ such that we can mix $\psi$ and $\bar{\psi}$ to obtain a mixed state that can also be obtained by mixing $\phi_1$ and $\phi_2$.
+Geometrically, consider a Bloch sphere.
+If $\psi$ is a superposition of $\phi_1$ and $\phi_2$ then all three states are on the surface of the same sphere.
+The chord connecting $\phi_1$ and $\phi_2$ represents the space of all possible mixtures of $\phi_1$ and $\phi_2$.
+Now, take a point $\rho$ on that chord, and consider the line that passes through $\rho$ and $\psi$.
+The same line will traverse the sphere at another point, which we call $\bar{\psi}$.
+Since $\rho$ is on the chord that connects $\psi$ and  $\bar{\psi}$, it is a mixture of the two.
+Therefore the vector space structure can be recovered by knowing how pure state mix.
+
+Note that the fact that the same statistical mixture can be decomposes in multiple ways is a characteristic feature of quantum mechanics, that is not present in classical mechanics.
+Using the same characterization, then, we consistently conclude that in classical mechanics there are no superpositions precisely because, given a pure state $\psi$, we can never find a corresponding state $\bar{\psi}$.
+
+The argument, then, goes like this.
+On physical grounds, we must be able to construct mixture.
+On physical grounds, we must be able to tell whether the same mixture can be prepared in different ways.
+The vector space structure captures exactly which pairs of pure states produce indistinguishable mixtures.
+The vector space structure exactly captures that structure and therefore it is fully physically justified.
+\textbf{The vector space structure is physical.}
+
+\subsection{Inner product}
+
+The inner product defines a geometrical structure on top of the vector space.
+Namely, it gives an idea of size of each vector and of angle between vectors.
+In quantum mechanics, the square of the inner product
+\begin{equation}
+	p(\psi_1, \psi_2) = |\<\psi_1| \psi_2\>|^2.
+\end{equation}
+represent the probability of measuring one state having prepared another.
+One may already accept this as physical, as these probabilities define all observable features of quantum mechanics.
+There is, again, a difficulty.
+
+Understanding the inner product as giving the probability of measurements also brings in one of the most controversial aspects of the theory: measurements.
+This is, again, a problem since we want our reasoning to be independent of interpretation.
+Another issue is that, technically, the inner product is linked to decomposition in linear combinations, to superposition, which are done at equal time.
+The probability above, instead, is a probability of transition.
+
+Given that we have characterized superposition in terms of mixtures, we should see whether the inner product characterizes something fundamental about them, and indeed it does.
+Given two pure states, the entropy of the mixture is a strictly monotonic, and therefore invertible, function of the square of the inner product.
+The greatest entropy is given by mixing orthogonal states, which gradually decreases as the angle in the inner product space decreases.
+We can then understand the geometrical structure of the pure quantum states as an entropic structure, much like it happens for classical mechanics.\footnote{In classical mechanics, the Liouville measure returns the count of states within a region, which is used in the fundamental postulate of statistical mechanics.
+The symplectic form gives us the Liouville measure for each independent degree of freedom.}
+
+The argument, then, goes like this.
+On physical grounds, for each statistical mixture, the thermodynamic entropy must be defined.
+Therefore it must be defined for equal mixtures of two pure states as well.
+The inner product exactly captures how that quantity combines and therefore it is fully physically justified.
+\textbf{The inner product structure is physical.}
+
+\subsection{Physicality of inner product space structure}
+
+First we want to show that the vector space structure is not, by itself, physical.
+Since quantum states are represented by rays of a Hilbert space, and since rays can be defined simply on top of the vector space structure alone, one may think that the vector space structure, by itself, is physical.
+The issue is that, without a norm, one cannot make physical sense of the coefficients one uses for linear combinations, therefore superpositions becomes ill-defined.
+
+The following proposition tells us that a vector representing a state can be written as a linear combination of vectors of other two states with arbitrary coefficients.
+\begin{prop}\label{vector_insufficient}[The vector space structure is not enough to uniquely identify superpositions.] Let $\mathcal{H}$ be a complex vector space and let $\textbf{P}(\mathcal{H})$ be its projective space.
+Let $\psi, \phi_1, \phi_2 \in \mathcal{H}$ such that $\psi = c_1 \phi_1 + c_2 \phi_2$ for some $c_1, c_2 \in \mathbb{C}$.
+Let $\pj{\psi}, \pj{\phi_1}, \pj{\phi_2} \in \textbf{P}(\mathcal{H})$ be the corresponding rays in the projective space.
+Then for any $\hat{c}_1, \hat{c}_2 \in \mathbb{C}$ we can find $\hat{\phi}_1 \in \pj{\phi_1}$ and $\hat{\phi}_2 \in \pj{\phi_2}$ such that
+$$\psi = \hat{c}_1 \hat{\phi}_1 + \hat{c}_2 \hat{\phi}_2.$$
+\end{prop}
+
+\begin{proof}
+Let $\hat{\phi}_1 = \frac{c_1}{\hat{c}_1} \phi_1$.
+Since they differ only by a constant, they are elements of the same ray.
+That is, $\hat{\phi}_1 \in \pj{\phi_1}$.
+Similarly, we let $\hat{\phi}_2 = \frac{c_2}{\hat{c}_2} \phi_2$ and we have $\hat{\phi}_2 \in \pj{\phi_2}$.
+We have
+\begin{equation}
+\begin{aligned}
+	\hat{c}_1 \hat{\phi}_1 + \hat{c}_2 \hat{\phi}_2 &= \frac{c_1}{\hat{c}_1} \phi_1 + \frac{c_2}{\hat{c}_2} \phi_2 \\
+	&= c_1 \phi_1 + c_2 \phi_2 \\ 
+	&= \psi.
+	\end{aligned}
+\end{equation}
+This concludes the proof
+\end{proof}
+
+Given that the same state is identified by infinitely many vectors, we have infinitely many linear combinations with all possible coefficients.
+Without knowing the norm of the vectors, it is impossible to identify which coefficients represent actual probability amplitudes: the square of norm of the coefficient will represent probability only in that case.
+The vector space structure by itself is therefore insufficient to provide a clear connection to physics.
+
+To argue that the inner product space structure, as a whole, is physical, one could argue that the structure characterizes superpositions, since it provides a norm, and the Born rule.
+These are features required by quantum mechanics, and therefore necessary and physical motivated.
+Many physicists will already find this argument convincing enough, however we believe it has a few problems.
+
+First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs.
+This is an area where different interpretations may differ,\footnote{TODO: add citations} and whether a mathematical representation is physical should not be interpretation dependent.
+Second, the Born rule, in its standard characterization, brings in one of the most controversial aspects of the theory: measurements.
+This is also suffers from an interpretation problem.
+Lastly, the inner product is linked to both the Born rule, which is a probability of transition, and linear decomposition into superposition, which is done at equal time.
+The goal, then, is a characterization of our physical requirement that is more necessary, more cohesive and interpretation independent.
+
+We will do this by characterizes the space of statistical mixtures, which are represented mathematically by the space of density operators.
+To have a physical sense of how this works, let us rephrase the above two requirements in terms of properties of mixtures.
+We show that a state is a linear superposition of
+
+The first is that it is not clear exactly what superpositions are.
+Some may argue that superpositions are just a mathematically convenient representation of the state, and express the potentiality of transition during measurement, not an actuality of the state itself.
+The second is
+
+\begin{prop}\label{prop_densitySpan}
+	Let $\mathcal{H}$ be a complex inner product space.
+Let $\{\psi_i\}, \{\phi_j\} \in \mathcal{H}$ be two finite sets of normalized vectors.
+Suppose there exists a density operator that can be expressed as a strict convex combination of either set, that is we can find two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}^+$ such that
+	\begin{enumerate}
+		\item $\sum_i a_i = 1$
+		\item $\sum_j b_j = 1$
+		\item $\sum_i a_i |\psi_i\>\<\psi_i| = \sum_j b_j |\phi_j\>\<\phi_j|$,
+	\end{enumerate}
+	then the span of the two sets is the same.
+\end{prop}
+
+\begin{proof}
+	Consider $\rho = \sum_i a_i |\psi_i\>\<\psi_i|$. For a unitary vector $\varphi \in \mathcal{H}$ we have
+	$$p(\varphi|\rho)=tr(\rho |\varphi\>\<\varphi|) =  \sum_i a_i |\<\psi_i|\varphi\>|^2$$
+	Since all $a_i$ are positive, $p(\varphi|\rho) = 0$ if and only if $\varphi$ is orthogonal to all $\psi_i$.
+
+	Now suppose $\rho = \sum_i a_i |\psi_i\>\<\psi_i| = \sum_j b_j |\phi_j\>\<\phi_j|$.
+Then a vector $\varphi$ is orthogonal to all $\psi_i$ if and only if it is also orthogonal to all $\phi_i$, which means $span(\{\psi_i\}) = span(\{\phi_j\})$) as they have the same orthogonal complement.
+\end{proof}
+
+\begin{prop}\label{prop_decomposition}
+	Let $\mathcal{H}$ be a complex inner product space.
+Let $\psi \in \mathcal{H}$ be a normalized state and $\rho : \mathcal{H} \to \mathcal{H}$ a density operator of finite rank such that $\<\psi|\rho|\psi\>\neq 0$.
+Then we can express $\rho$ as a mixture of $\psi$.
+That is, there exists a finite set of normalized states $\{\psi_i\} \in \mathcal{H}$ such that $\psi_1 = \psi$ and $\rho = \sum_i p_i |\psi_i\>\<\psi_i|$.
+\end{prop}
+
+\begin{proof}
+	The set of all possible density operators is a convex set, which can be understood as a subset of a vector space $V$.
+Let $\rho$ be a density operator such that $\<\psi|\rho|\psi\>\neq 0$.
+If $\rho = |\psi\>\<\psi|$, we are done.
+If not, the two elements will identify a line, which must intersect the boundary of the convex set in two places.
+One must be $\psi$, since $\psi$ is a pure state and therefore an extreme point.
+Let $\rho_1$ be the other point.
+If it is an extreme point, we are done as $\rho$ is an affine combination of $\psi$ and $\rho_1$.
+If not, the boundary of the convex set in a neighborhood of $rho_1$ is not strictly convex, and therefore is a subspace $V_1$ of the vector space $V$.
+Note that $V_1$ does not include $\psi$ and it does not span the direction identified by $\rho_1$ and $\psi$.
+Furthermore, we can't have $\rho_1 = \rho$ as $\<\psi|\rho|\psi\>\neq 0$.
+Therefore we have that $\rho$ is an affine combination of $\psi$ and $\rho_1$, which must be a density operator of finite rank lower than $\rho$.
+We can express $\rho_1$ as a convex combination of finitely many extreme points $\psi_2, \psi_3, ..., \psi_n$, each of them corresponding to a pure state.
+We have that $\rho$ is an affine combination of pure states that include $\psi$.
+\end{proof}
+
+\begin{prop}[Superposition is multiple decomposition of mixed states.]
 Let $\mathcal{H}$ be a complex vector space.
-Let $\psi, \phi_1, \phi_2 \in \mathcal{H}$ be pure states.
-Then $\psi$ is a superposition of $\phi_1$ and $\phi_2$, that is $\psi = c_1 \phi_1 + c_2 \phi_2$, if and only if there is another state $\bar{\psi}$ such that an equal mixture of $\psi$ and $\bar{\psi}$ is also a mixture of $\phi_1$ and $\phi_2$.
-That is
-\begin{equation}
-	\frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\bar{\psi}\>\<\bar{\psi}| = p_1|\phi_1\>\<\phi_1| + p_2|\phi_2\>\<\phi_2|
-\end{equation}
+Let $\psi \in \mathcal{H}$ be a normalized vector and $\{\phi_j\} \in \mathcal{H}$ a finte sets of normalized vectors that does not contain $\phi$.
+Then $\psi$ is a superposition of $\{\phi_j\}$ if and only if there exists a density operator $\rho$ that can be expressed both as a strict convex combination of $\{\phi_i\}$ and of another set of normalized vectors that contain $\psi$.
+That is, we can find a countable set of normalized vectors $\{\psi_i\} \in \mathcal{H}$ and two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}^+$ such that
+\begin{enumerate}
+	\item $\psi_1 = \psi$
+	\item $\sum_i a_i = 1$
+	\item $\sum_j b_j = 1$
+	\item $\sum_i a_i |\psi_i\>\<\psi_i| = \sum_j b_j |\phi_j\>\<\phi_j|$,
+\end{enumerate}
+\end{prop}
 
+\begin{proof}
+	Suppose $\psi$ is a normalized state that can be written as $\psi = \sum_{i=1}^n c_i \phi_i$ where $\{\phi_i\} \in \mathcal{H}$ a finte sets of normalized vectors that do not contain $\psi$.
+Since $\{\phi_j\}$ are not necessarily orthogonal, it is not necessarily true that $|\<\psi|\phi_i\>| = |c_i|^2 \neq 0$.
+However, since $\psi$ is normalized, there must be at least two vectors in $\{\phi_i\}$ such that $|\<\psi|\phi_i\>| \neq 0$.
+We can then find $\{p_i\} \in \mathbb{R}+$ for which $\sum_i p_i =1$, such that $\rho = \sum_i p_i |\phi_j\>\<\phi_j|$ and $\<\psi|\rho|\psi\> \neq 0$.
+By Proposition \ref{prop_decomposition}, $\rho$ can be expressed as a convex combination of a set of normalized states that include $\psi$.
+
+	Conversely, suppose we have a density operator $\rho$ that can be expressed both as a strict convex combination of $\{\phi_i\}$ and of another set of normalized vectors that contain $\psi$.
+By Proposition \ref{prop_densitySpan} the span of $\{\phi_i\}$ will include $\psi$ and therefore $\psi$ is a superposition of $\{\phi_i\}$.
+\end{proof}
+
+We can now show that the inner product space structure is physical in the sense that it is exactly the structure that is needed to model quantum mixed states and their entropy.
+One key difference between classical and quantum mechanics is that there is no single decomposition of quantum mixed states in terms of pure states.
+We can show that relationships of superposition are equivalent to relationships of multiple decomposition.
+
+\begin{proof}
+	Let's first address the trivial case in which $\phi_1 = \phi_2$.
+Suppose that $\psi = c_1 \phi_1 + c_2 \phi_2$, then $\psi = \phi_1$.
+We set $\bar{\psi} = \psi$ and find $p_1|\psi\>\<\psi| + p_2|\bar{\psi}\>\<\bar{\psi}| = \frac{1}{2}|\phi_1\>\<\phi_1| + \frac{1}{2}|\phi_2\>\<\phi_2|$ for any $p_1, p_2 \in \mathbb{R}$ such that $p_1+p_2=1$.
+Conversely, suppose $p_1|\psi\>\<\psi| + p_2|\bar{\psi}\>\<\bar{\psi}| = \frac{1}{2}|\phi_1\>\<\phi_1| + \frac{1}{2}|\phi_2\>\<\phi_2|$ for some $p_1, p_2 \in \mathbb{R}$ such that $p_1+p_2=1$.
+Then $p_1|\psi\>\<\psi| + p_2|\bar{\psi}\>\<\bar{\psi}| = |\phi_1\>\<\phi_1|$ which can be true only if $\psi=\bar{\psi}=\phi_1$.
+
+	Now let's address the case in which $\phi_1 \neq \phi_2$.
+Suppose that $\psi = c_1 \phi_1 + c_2 \phi_2$.
+Then $\psi$ is part of the subspace defined by $\phi_1$ and $\phi_2$, which means all the states can be represented as points on the surface of a Bloch sphere.
+The mixed state $\rho=\frac{1}{2}|\phi_1\>\<\phi_1| + \frac{1}{2}|\phi_2\>\<\phi_2|$ is the midpoint of the chord connecting $\phi_1$ and $\phi_2$.
+Now, take the line that connects $\rho$ with $\psi$.
+This line will traverse the sphere in another point $\hat{\psi}$.
+As $\rho$ sits on the chord between $\psi$ and $\bar{\psi}$, it is a mixture of the two states for some $p_1$ and $p_2$.
+Conversely, suppose $\rho = p_1|\psi\>\<\psi| + p_2|\bar{\psi}\>\<\bar{\psi}| = \frac{1}{2}|\phi_1\>\<\phi_1| + \frac{1}{2}|\phi_2\>\<\phi_2|$.
+This means that the subspace spanned by $\psi$ and $\bar{\psi}$ is the same as the one spanned by $\phi_1$ and $\phi_2$, which means that $\psi = c_1 \phi_1 + c_2 \phi_2$ for some $c_1, c_2 \in \mathbb{C}$ such that $|p_1|^2+|p_2|^2=1$.
+\end{proof}
+
+The norm of the inner product can be shown to be equivalent to the entropy of uniform distributions over state pairs.
+This gives us a geometric understanding of the inner product at equal time, instead of a probability of transition.
+
+\begin{prop}[Inner product is an entropic structure]
 Let $\mathcal{H}$ be a complex Hilbert space. Let $I : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the von Neumann entropy of the equal mixture of two pure states. That is
@@
 \end{equation}
+\end{prop}
+
+\begin{proof}
+	Copy proof from other paper.
+\end{proof}
 

### 2023-06-20 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Reviewed main text and appendix for inner product structure

@@
 The set of countable sequences of zeros and one can be understood as the binary expansion of real numbers between zero and one, which has cardinality of the continuum.
-Therefore, any definition that would require a set of physical objects to have cardinality greater than the continuum, then, is \textbf{unphysical}.\footnote{For example, non separable Hilbert spaces have cardinality greater than the continuum.}
+Therefore, any definition that would require a set of physical objects to have cardinality greater than the continuum, then, is \textbf{unphysical}.
+For example, non separable Hilbert spaces have cardinality greater than the continuum, and therefore should be considered unphysical.
 
@@
 In general, testing physicality is harder as we need to check that all the mathematical features map to physical feature and no physical feature is left behind.
-Clearly, postulating that ``the state space of a quantum states has cardinality no grater than that of the continuum'' is certainly \textbf{physical}.
-We now need to check the rest.
+Clearly, postulating that ``the state space of a quantum states has cardinality no grater than that of the continuum'' is certainly \textbf{physical} but, however, this tells us little.
+Let us then see what elements of Hilbert spaces should we consider physical in the context of quantum mechanics.
 
@@
 That is, it is a vector product, with an inner product, closed under Cauchy sequences.
-Let us look at the first two structure briefly, as they are straight forward.
-We will leave more detail in the appendix.
+We will look at the first two structures briefly, leaving all detail to the appendix, as their physical significance is not surprising.
 
@@
 If one regards superpositions as physical requirements of quantum theory, one may think that this structure is physical.
-However, the ability to interpret linear combination as superpositions requires the ability to keep track of normalization.
-If we can only know whether two vectors are in the same ray, without knowing which ones are normalized, we are not able to recover the idea of a probability amplitude, as those are the coeffcients of a linear combination between normalized states.
+However, the ability to interpret linear combination as superpositions requires the ability to keep track of orthonormal vectors.
+If we can only know whether two vectors are in the same ray, we are not able to recover the idea of a probability amplitude, as those are the coefficients of a linear combination between normalized and orthogonal states that return another normalized state.
 Therefore the vector space structure, by itself, is not physical.
 
-If we add the inner product, we can define the norm of vectors and therefore we can indeed express superpositions.
-The inner product structure also defines probabilities of transitions.
+If we add the inner product, we can define norm and orthogonality and therefore we can indeed express superpositions.
+Additionally, the inner product structure also defines the Born rule, the transition probability during measurement.
 If one is comfortable in considering superpositions and the Born rule to be physical requirements of quantum mechanics, then the inner product structure is \textbf{physical}.
-However, some may not consider superpositions as actual physical, but more as mathematical constructs; others may object that the Born rule, in its standard interpretation, defines probability of transitions while a state space should be defined at equal time; others can argue that the states are defined by the rays, and therefore the requirements should only be in terms of those objects.
-For those, we have an alternative, more necessary justification, which we include in the appendix, and can be summarized as follows.
+However, some may not consider superpositions as actual physical objects, but more as mathematical constructs; others may object that the Born rule, in its standard interpretation, defines probability of transitions while a state space should be defined at equal time; others can argue that the states are defined by the rays, and therefore the requirements should only be in terms of those objects.
+For those and other reasons, we have an alternative, more necessary justification, which we include in the appendix, and can be summarized as follows.
 
@@
 Mathematically, these form a convex set, that physically represents a set that allows statistical mixtures.
-The geometry of the space of statistical ensembles is uniquely defined by the inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtues.\footnote{There is a rich literature on Generalized Probability Theory in which people reconstruct Quantum Theory by imposing requirements on convex spaces.} For example, we can characterize superposition in the following way: a pure state $\psi$ is a superposition of two other pure states $\phi_1$ and $\phi_2$ if and only if there exists another pure states $\bar{\psi}$ such that we can mix $\psi$ and $\bar{\psi}$ to obtain a mixed state that can also be obtained by mixing $\phi_1$ and $\phi_2$.
-That is, the ability to create superposition is conceptually equivalent to the ability to prepare the same mixture in different ways.
-The inner product can be shown to be equivalent to determining the entropy of all equal mixtures of state pairs.
+The geometry of the space of statistical ensembles is uniquely defined by the inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtues.\footnote{There is a rich literature on Generalized Probability Theory in which people reconstruct Quantum Theory by imposing requirements on convex spaces.} For example, we can characterize superposition in the following way: a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ if and only if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
+That is, the ability to create superposition is conceptually equivalent to the ability to prepare the same mixture in different ways, another feature particular to quantum mechanics.
+Additionally, the inner product can be shown to be equivalent to determining the entropy of all equal mixtures of state pairs.
+These requirements are necessary in the sense that both classical and quantum mechanics need to keep track how ensemble combine under mixing and what entropy each ensemble have.
 
-If we require, then, that the theory reproduces the properties of quantum mixtures, how they can be decomposed in terms of pure states and what their entropy is, the inner product space is recovered.
+If we require, then, that the theory reproduces the properties of quantum mixtures, how they can be decomposed in terms of pure states and what their entropy is, the complex inner product space is recovered as a structure that recovers the correct properties.
 As we said, a physical theory must tell us how to properly characterized statistical mixture, therefore the requirement is necessitated by the physics.
 \textbf{The inner product space structure is physical}.
-
-----
-
- Geometrically, consider a Bloch sphere.
-If $\psi$ is a superposition of $\phi_1$ and $\phi_2$ then all three states are on the surface of the same sphere.
-The chord connecting $\phi_1$ and $\phi_2$ represents the space of all possible mixtures of $\phi_1$ and $\phi_2$.
-Now, take a point $\rho$ on that chord, and consider the line that passes through $\rho$ and $\psi$.
-The same line will traverse the sphere at another point, which we call $\bar{\psi}$.
-Since $\rho$ is on the chord that connects $\psi$ and  $\bar{\psi}$, it is a mixture of the two.
-Therefore the vector space structure can be recovered by knowing how pure state mix.
-
-Note that the fact that the same statistical mixture can be decomposes in multiple ways is a characteristic feature of quantum mechanics, that is not present in classical mechanics.
-Using the same characterization, then, we consistently conclude that in classical mechanics there are no superpositions precisely because, given a pure state $\psi$, we can never find a corresponding state $\bar{\psi}$.
-
-The argument, then, goes like this.
-On physical grounds, we must be able to construct mixture.
-On physical grounds, we must be able to tell whether the same mixture can be prepared in different ways.
-The vector space structure captures exactly which pairs of pure states produce indistinguishable mixtures.
-The vector space structure exactly captures that structure and therefore it is fully physically justified.
-\textbf{The vector space structure is physical.}
-
-\subsection{Inner product}
-
-The inner product defines a geometrical structure on top of the vector space.
-Namely, it gives an idea of size of each vector and of angle between vectors.
-In quantum mechanics, the square of the inner product
-\begin{equation}
-	p(\psi_1, \psi_2) = |\<\psi_1| \psi_2\>|^2.
-\end{equation}
-represent the probability of measuring one state having prepared another.
-One may already accept this as physical, as these probabilities define all observable features of quantum mechanics.
-There is, again, a difficulty.
-
-Understanding the inner product as giving the probability of measurements also brings in one of the most controversial aspects of the theory: measurements.
-This is, again, a problem since we want our reasoning to be independent of interpretation.
-Another issue is that, technically, the inner product is linked to decomposition in linear combinations, to superposition, which are done at equal time.
-The probability above, instead, is a probability of transition.
-
-Given that we have characterized superposition in terms of mixtures, we should see whether the inner product characterizes something fundamental about them, and indeed it does.
-Given two pure states, the entropy of the mixture is a strictly monotonic, and therefore invertible, function of the square of the inner product.
-The greatest entropy is given by mixing orthogonal states, which gradually decreases as the angle in the inner product space decreases.
-We can then understand the geometrical structure of the pure quantum states as an entropic structure, much like it happens for classical mechanics.\footnote{In classical mechanics, the Liouville measure returns the count of states within a region, which is used in the fundamental postulate of statistical mechanics.
-The symplectic form gives us the Liouville measure for each independent degree of freedom.}
-
-The argument, then, goes like this.
-On physical grounds, for each statistical mixture, the thermodynamic entropy must be defined.
-Therefore it must be defined for equal mixtures of two pure states as well.
-The inner product exactly captures how that quantity combines and therefore it is fully physically justified.
-\textbf{The inner product structure is physical.}
 
@@
 Since quantum states are represented by rays of a Hilbert space, and since rays can be defined simply on top of the vector space structure alone, one may think that the vector space structure, by itself, is physical.
-The issue is that, without a norm, one cannot make physical sense of the coefficients one uses for linear combinations, therefore superpositions becomes ill-defined.
+The issue is that, without an inner product, one cannot make physical sense of the coefficients one uses for linear combinations, therefore superpositions becomes ill-defined.
 
-The following proposition tells us that a vector representing a state can be written as a linear combination of vectors of other two states with arbitrary coefficients.
+The following proposition tells us that there are infinitely many ways to decompose a vector representing a state in terms of vectors that represent two other states.
 \begin{prop}\label{vector_insufficient}[The vector space structure is not enough to uniquely identify superpositions.] Let $\mathcal{H}$ be a complex vector space and let $\textbf{P}(\mathcal{H})$ be its projective space.
@@
 \end{equation}
-This concludes the proof
+This concludes the proof.
 \end{proof}
 
-Given that the same state is identified by infinitely many vectors, we have infinitely many linear combinations with all possible coefficients.
-Without knowing the norm of the vectors, it is impossible to identify which coefficients represent actual probability amplitudes: the square of norm of the coefficient will represent probability only in that case.
+The intuition that a state has a single linear decomposition in terms of other vectors requires, first of all, that the vectors are normalized.
+To identify the coefficient as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires the decomposition to be done on orthogonal vectors.
+That is, linear combinations require the inner product to be understood physically as superpositions.
 The vector space structure by itself is therefore insufficient to provide a clear connection to physics.
 
-To argue that the inner product space structure, as a whole, is physical, one could argue that the structure characterizes superpositions, since it provides a norm, and the Born rule.
-These are features required by quantum mechanics, and therefore necessary and physical motivated.
-Many physicists will already find this argument convincing enough, however we believe it has a few problems.
+We could now use the above reasoning to argue that the inner product space structure, as a whole, is physical, because it indeed characterizes both superpositions and the Born rule.
+These are features required by quantum mechanics, and therefore necessary and physically motivated.
+Many will find this argument convincing enough, however there a couple of objective problems.
 
 First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs.
-This is an area where different interpretations may differ,\footnote{TODO: add citations} and whether a mathematical representation is physical should not be interpretation dependent.
-Second, the Born rule, in its standard characterization, brings in one of the most controversial aspects of the theory: measurements.
-This is also suffers from an interpretation problem.
-Lastly, the inner product is linked to both the Born rule, which is a probability of transition, and linear decomposition into superposition, which is done at equal time.
-The goal, then, is a characterization of our physical requirement that is more necessary, more cohesive and interpretation independent.
+This is an area where different interpretations may differ,\footnote{TODO: add citations} some arguing that they are just mathematical representations, and the physicality of a mathematical representation should not be interpretation dependent.
+Second, the Born rule, in its standard characterization, brings in one of the most controversial aspects of the theory, measurements, which suffers from an interpretation problem as well.
+Lastly, the inner product is critical for both the Born rule, which is a transition probability of transition before and after the measurement, and linear decomposition into superposition, which is done at equal time.
+The goal, then, is to find a characterization of our physical requirement that is interpretation independent, more cohesive and perhaps more physically necessary.
 
-We will do this by characterizes the space of statistical mixtures, which are represented mathematically by the space of density operators.
-To have a physical sense of how this works, let us rephrase the above two requirements in terms of properties of mixtures.
-We show that a state is a linear superposition of
+The idea is to reformulate the requirements that lead to superposition and the Born rules in terms of equivalent requirements of statistical mixtures.
+Mathematically, these will be in terms of the space of density operators in stead of the vector space itself.\footnote{Similar efforts can be found in the literature of Quantum Logic and of Generalized Probability Theories.
+TODO: add citations}.
 
-The first is that it is not clear exactly what superpositions are.
-Some may argue that superpositions are just a mathematically convenient representation of the state, and express the potentiality of transition during measurement, not an actuality of the state itself.
-The second is
+The idea is to link superposition to another feature of quantum mechanics: the non-unique decomposition of mixed state into pure states.
+For example, if we take the maximally mixed state for a spin 1/2 system, this can be implemented with an equal mixture of spin up and spin down, or with an equal mixture of spin left and spin right.
+The idea is that a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ exactly if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
+To keep the reasoning as general as possible, we will not require $\{\phi_i\}$ to be orthogonal to each other.
 
@@
 
-\begin{prop}[Superposition is multiple decomposition of mixed states.]
+\begin{prop}\label{prop_superpositionIsDecomposition}[Superposition is multiple decomposition of mixed states.]
 Let $\mathcal{H}$ be a complex vector space.
@@
 
-The norm of the inner product can be shown to be equivalent to the entropy of uniform distributions over state pairs.
+As for the inner product, we show that it can be fully characterized, instead of by the Born rule, but the entropy of equal mixtures over state pairs.
 This gives us a geometric understanding of the inner product at equal time, instead of a probability of transition.
 
-\begin{prop}[Inner product is an entropic structure]
+\begin{prop}\label{prop_innerProductIsEntropy}[Inner product is an entropic structure]
 Let $\mathcal{H}$ be a complex Hilbert space. Let $I : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the von Neumann entropy of the equal mixture of two pure states. That is
@@
 
+We are then able to argue for the following
+\begin{prop}
+	The description of quantum states as rays of a complex inner product space is physically justified.
+\end{prop}
+\begin{justification}
+	Since physical preparations and measurements are most of the type not perfect, a physical theory must define statistical ensembles.
+Part of that definition includes keeping track statistical ensembles under statistical mixtures, that is which particular mixture is obtained by mixing particular states in particular proportions.
+A particular feature of quantum mixed states is that, unlike classical states, they do not in general provide a unique decomposition in terms of pure states.
+This non-unique decomposition, as shown by proposition \ref{prop_superpositionIsDecomposition}, is equivalent to the ability to create superpositions of pure state.
+The inner product structure is therefore the exact structure needed to characterize quantum ensembles and their properties, which is required by the physics.
+
+	Additionally, a physical theory must be able to define the entropy of each ensemble, as required by thermodynamics.
+In quantum mechanics, the thermodynamic entropy is recovered by the von Neumann entropy, which Proposition \ref{prop_innerProductIsEntropy} show it is exactly captured by the inner product.
+This again shows that the inner product structure is required to capture a necessary physical requirement.
+\end{justification}
+
 \end{document}

### 2023-06-20 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Minor edits

@@
 The intuition that a state has a single linear decomposition in terms of other vectors requires, first of all, that the vectors are normalized.
-To identify the coefficient as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires the decomposition to be done on orthogonal vectors.
-That is, linear combinations require the inner product to be understood physically as superpositions.
+To identify the coefficient as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires the decomposition to be done on orthogonal vectors.\footnote{TODO: mention Gram-Schmidt orthonormalization} That is, linear combinations require the inner product to be understood physically as superpositions.
 The vector space structure by itself is therefore insufficient to provide a clear connection to physics.
@@
 Let $\mathcal{H}$ be a complex vector space.
-Let $\psi \in \mathcal{H}$ be a normalized vector and $\{\phi_j\} \in \mathcal{H}$ a finte sets of normalized vectors that does not contain $\phi$.
+Let $\psi \in \mathcal{H}$ be a normalized vector and $\{\phi_j\} \in \mathcal{H}$ a finte sets of normalized vectors that does not contain $\psi$.
 Then $\psi$ is a superposition of $\{\phi_j\}$ if and only if there exists a density operator $\rho$ that can be expressed both as a strict convex combination of $\{\phi_i\}$ and of another set of normalized vectors that contain $\psi$.

### 2023-06-20 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Updated introduction

@@
 
-Von Neumann's \textit{Mathematische Grundlagen der Quantenmechanik} (\citeyear{von_neumann_mathematische_1996}) crowned Hilbert spaces as (i) the unifying framework for the states in Schrödinger's wave mechanics and Heisenberg-Born-Jordan's matrix mechanics that is (ii) mathematically rigorous.
-Arthur Wightman, one of the people who followed von Neumann's aims to unify and axiomatize quantum field theory (\citeyear{wightman_quantum_1956}), took his book to be the fulfillment of a project Hilbert himself kicked off with his problems on the foundations of mathematics (\citeyear[p.
+Hilbert spaces have been the cornerstone of quantum mechanics since von Neumann (\cite{von_neumann_mathematische_1996}).
+Their use made available a vast number of theorems (Gleason, Riesz, Wigner, Stone, etc.) that made up the mathematical backbone of the theory and also represented a role model of mathematical rigor that people working on quantum field theory would try to achieve decades later \cite{haag_discussion_2010}.
+Wightman even took von Neumann's book to be the fulfillment of a project Hilbert himself kicked off with his problems on the foundations of mathematics (\cite[p.
 157]{browder_hilberts_1976}): to axiomatize the physical sciences.
-Despite the undeniable importance of wanting to set physics on a firm mathematical footing, we must remember that by ``axioms,'' they meant ``physically motivated constraints'' more than statements that can be immune to question, rethink, and, if necessary, revise.\footnote{For a more extended discussion of von Neumann's axiomatics and their physical inputs, see \citep{redei_soft_2006}.}\\
+But, despite the undeniable importance of wanting to set physics on a firm mathematical footing, we must remember that by ``axioms,'' they meant ``physically motivated constraints'' more than statements that can be immune to question, rethink, and, if necessary, revise.\footnote{For a more extended discussion of von Neumann's axiomatics and their physical inputs, see \cite{redei_soft_2006}.}
 
-In this letter, we focus on an assumption that poses an array of interesting questions: that quantum states are vectors in Hilbert spaces.
-At first glance, some of their shortcomings (states are rays, not vectors; the Dirac delta is a distribution, not a function; some elements from $L^2(\mathbb{R})$ don't seem apt candidates to be wave functions) are well-known, but deemed unproblematic.
-Additionally, a handful of theorems (Gleason, Riesz, Wigner, Stone, Stone-von Neumann, Wigner, etc.) rely heavily on the structure of a Hilbert space.
-While these results are rarely mentioned when students are first introduced to quantum mechanics, they are crucial to substantiate the claim that Hilbert spaces are the right structure to characterize quantum states.
-Therefore, it's not a surprise that Hilbert spaces have become fossilized as parts of axiomatic systems\footnote{Algebraic quantum field theory (see, for instance, \citep{haag_local_1996}) inverts the narrative of ``quantum states are normed vectors in a Hilbert space, and observables are bounded, self-adjoint operators acting on that space'' to ``canonical (anti)commutation relations of observables generate an algebra, and states are assignments of expectation values to elements of the algebra.'' As such, it doesn't \textit{need} to talk about Hilbert spaces to do quantum mechanics, but it has resources, like the GNS construction, that allow their advocates to say that, since they're recovering Hilbert spaces from algebras of observables, what they're doing ends up being similar to the quantum mechanics that's more familiar to physicists.} without even considering whether we can do quantum mechanics with weaker assumptions.
+Now, there are well-known shortcomings of Hilbert spaces, but they are deemed unproblematic: states are rays, not vectors; the Dirac delta is a distribution, not a function; and some elements from $L^2(\mathbb{R})$ don't seem apt candidates to be wave functions.
+Given the importance Hilbert spaces have had and the issues we have let slide, it's not a surprise that they have become fossilized as parts of axiomatic systems\footnote{Even axiomatic formulations of quantum field theory (see e.
+g.
+\cite{haag_local_1996}) that don't need to use Hilbert spaces to make meaningful claims about quantum-mechanical observables and states can reassure physicists unfamiliar with this approach by recovering the Hilbert spaces they know and love through the GNS construction.} without even considering whether we can do quantum mechanics with weaker assumptions.
 So, despite all their success, perhaps it's time to ask if using Hilbert spaces to describe quantum states, an assumption whose physical motivation was meant to be revised as time passed, has become a dogma.

### 2023-06-21 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Reviewed main text

@@
 
-If we require, then, that the theory reproduces the properties of quantum mixtures, how they can be decomposed in terms of pure states and what their entropy is, the complex inner product space is recovered as a structure that recovers the correct properties.
-As we said, a physical theory must tell us how to properly characterized statistical mixture, therefore the requirement is necessitated by the physics.
+The argument can be summed up as follows.
+On physical grounds, for any physical theory, we must be able to prepare statistical mixtures and characterize their entropy.
+The complex inner product space structure is exactly required to model the correct relationships of quantum mixtures, therefore the mathematical definition is necessitated by physical requirements.
 \textbf{The inner product space structure is physical}.
@@
 
-Completeness of a Hilbert space tells us that the space includes the limit of every Cauchy sequence.
-This property is more difficult to characterize physically.
+We now turn to the last property of Hilbert spaces: completeness.
+Mathematically, this means that the space includes the limit of every Cauchy sequence; physically, this property is more difficult to characterize.
 We can look at an equivalent characterization:\footnote{Roman 2008, p.
@@
 the series converges in the space.
-This is now in terms of linear combinations, and therefore it is essentially telling us that we can perform a linear superposition if the sum of the norm of the coefficient converge.
-We are basically extending superposition to infinite sums.
+This is now in terms of linear combinations, and therefore it is essentially telling us that we can prepare states that are superposition of infinitely many elements or, equivalently as we have seen, mixed states comprised of infinitely many pure states.
 Is this a physical requirement of just mathematical convenience?
 
-As we saw before, superposition experimentally means splitting the state along multiple paths.
-The above mathematical expression means splitting the original state into infinitely many progressively small component, and be able to place them where we please.
-It also assumes, for example, that we can place them over an infinite spatial range.
-Since the limit must be in the space, we are literally saying that we can prepare states that are spread over infinite regions.
+Experimentally, we can imagining creating a superposition by splitting a beam along multiple paths, which is something typical, for example, in quantum optics.\footnote{TODO: cite} The above mathematical expression means splitting the original state into infinitely many progressively small components, and be able to place them where we please.
+However, it also assumes that we can place them over an infinite spatial range: we are literally saying that we can prepare states that are spread over infinite regions.
 This does not sound physical.
@@
 Clearly, this does not make physical sense.
-When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that observed value can be arbitrarily large, not that we can literally spread values over an infinite range.
+When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that observed value can be arbitrarily large, not that we can literally spread values over an infinite range.\footnote{This is the difference between potential infinity and actual infinity.}
 
-The argument, then, goes like this.
-On physical grounds, given a quantity that can take arbitrarily large values, like position, for each possible value there must exist some state whose expectation matches that value.
-On mathematical grounds, we require completeness, which forces us to posit states with infinite expectation value.
-But this, on physical grounds, is not teneable.
-Therefore \textbf{completeness in unphysical.}
+The argument can be summed up as follows.
+On physical grounds, we assume the existence of quantities, like position, that can take arbitrarily large values.
+This requires that for each possible value there must exist some state whose expectation matches that value.
+Requiring completeness mathematically forces us to posit states with an infinite expectation value, which cannot be implemented physically.
+The mathematical definition implicitly assumes objects that cannot be physically realized.
+\textbf{The use of completeness for the space of quantum systems in unphysical.}
 
@@
 
-We are forced to conclude that Hilbert spaces have unphysical requirements when applied to quantum mechanics.
-We clearly see that the issue is the closure of operation over infinite sequences, which is not surprising.\footnote{Large branches of mathematics were developed to handle infinities, both large and small.
-Weyl stated ``Mathematics is the science of the infinite, its goal the symbolic comprehension of the infinite with human, that is finite, means.'' Hermann Weyl, "The Open World: Three Lectures on the Metaphysical Implications of Science," (1932) as quoted in Mind and Nature: Selected Writings on Philosophy, Mathematics, and Physics (2009) ed.
-Peter Pesic.} Now we have to understand how big of a problem it is.
+It should not be surprising that closure over infinite sequences is problematic, as properly handling infinity is one of the key problems in mathematics.\footnote{Weyl stated ``Mathematics is the science of the infinite, its goal the symbolic comprehension of the infinite with human, that is finite, means.'' Hermann Weyl, "The Open World: Three Lectures on the Metaphysical Implications of Science," (1932) as quoted in Mind and Nature: Selected Writings on Philosophy, Mathematics, and Physics (2009) ed.
+Peter Pesic.
+Several branches of mathematics were developed to handle infinities, both large and small.
+} We should understand the impact on the physics of using completeness for that goal.
 
-On physical grounds, the state space of quantum system is the projective space of a complex inner product space.
-If the space is finite dimensional, the space is automatically complete.\footnote{This is exactly why, mathematically, completeness is nice as it makes infinite dimensional space behave like finite dimensional ones.} If not, we can always complete it by enlarging the space.
-In general, then, it is true that a quantum state is a ray of a Hilbert space, but the converse is not true: not all rays of the same Hilbert space will be valid quantum states.
+If an inner product space is finite dimensional, it is automatically complete.\footnote{This is exactly why, mathematically, completeness is nice as it makes infinite dimensional space behave like finite dimensional ones.} If the space is not complete, we can always complete it by enlarging the space.
+The prevalent attitude among physicists is that this pose no problem: we simply have a bigger set of objects than what we strictly need, and we simply discard what we do not need.
+Unfortunately, having a bigger space often creates bigger problems.
 
-This may not seem like a problem: we simply have a bigger set of object of what we strictly need.
-Unfortunately, in mathematics working on a bigger space is not always a good thing.
 For example, take the problem of defining volumes in three dimensional euclidean space $\mathbb{R}^3$.
@@
 Naively, we may want to define it on the set of all possible subsets.
-In this case, Banach–Tarski paradox tells us that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls of the same size.
-Clearly, this is a problem.
+In this case, the Banach–Tarski paradox tells us that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls of the same size.
+Clearly, this does not make physical sense.
 However, if we restrict ourselves with a measure on the Borel sets instead of all sets, we get a consistent definition.
-Therefore, in mathematics, bigger is not always better.\footnote{In another work, we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
-In many cases, we have found that physical requirements naturally map to well behaved spaces.}
+Therefore, in mathematics, bigger is not always better:\footnote{In another work, we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
+In many cases, we have found that physical requirements naturally map to well behaved spaces.} we need to look more closely at what the math dragged in.
 
-The above example is just a cautionary tale and it doesn't tell us what happens in this specific case.
-Therefore, let us give an example of the things that can go wrong.
 Unitary transformation play an important role in the theory of Hilbert spaces for at least two reasons.
@@
 The first distribution is a Gaussian, and therefore all moments of the distribution, the expectation for $x^n$, are finite.
-The second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all.
+The second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all moments above the first.
 We can now find a change of variable $y(x)$ that transforms one distribution into the other.
@@
 \end{equation}
-Therefore, thorough a variable change, we can change a state that has finite expectation value for all moments greater than the first into a state that has infinite expectation value for the same moments.
+Therefore, through a variable change, we can change a state that has finite expectation value for all moments greater than the first into a state that has infinite expectation value for the same moments.
 Variable changes, in terms of the Hilbert space, are unitary transformation.
@@
 
-Experimentally, these are not equivalent.
-Identifying a state through expectation value is a common technique, called quantum tomography.
-If a unitary transformation can change finite expectations to infinite ones, we are effectively changing what can be distinguished through a tomography.
+On physical grounds, these are clearly not equivalent.
+Identifying a state through expectation value is a common technique (i.e quantum tomography).
+If a unitary transformation can change finite expectations to infinite ones, we are effectively changing what can be distinguished experimentally.
 Not only we cannot measure infinite values, but different finite values would get lumped into the same infinite value.
-In the above example, if we increase the variance, all even moments would change, yet, after the transformation, they would all be infinite.
-Therefore the idea that such unitary transformation links to physically equivalent representations is, on physics ground, untenable.
+In the above example, if we increased the variance, one observer will see all the even moments change, yet the other would see them all equally infinite.
+The math, as it stands, lumps together physically inequivalent representations into the same equivalence class, making it impossible later to separate physical problem from unphysical ones.
 
@@
 \begin{equation}
-	y = \cos(\omega t) x + \sin(\omega t) \tan \left(\frac{\pi}{2}\erf(x)\right).
+	y(t) = \cos(\omega t) x + \sin(\omega t) \tan \left(\frac{\pi}{2}\erf(x)\right).
 \end{equation}
 This is clearly continuous in $t$ and, for $t=0$, we have the identity transformation and therefore, mathematically, there is no reason why this wouldn't be a valid time evolution.
-The effect is simply a stretching back and forth the wavefunctions, such that the state keeps oscillating between the above two functions.
-On physics grounds, however, we are going from finite to infinite expectations in finite time.
+What this does is stretching back and forth the wavefunction, such that the state keeps oscillating between the above two functions, making the expectation value oscillate from finite to infinite in finite time.
 This, again, is physically untenable.
 
-We hope that this gives an idea of the type of issues embedded in the mathematical definitions.
-When performing actual calculations, we do not see these issues because the physics that we are describing makes us work in a corner of the structure where things are reasonable.
-But when one tries to prove theorems or general results on quantum theory, one currently is forced to consider the entire Hilbert space.
-It is not inconceivable to expect that results that would be reasonable physically fail to valid when enlarging the space of possible states to those that have infinite expectation values and the space of physical transformations to those that map finite to infinite expectations.
-We therefore wonder whether some of the issues we have in quantum field theory (e.g.
-the inability to find a valid measure for path integrals, inequivalent representation of interacting theories (?), infinities) are simply a consequence of the use, at a fundamental level, of unphysical objects.
+Again, most of the time these issues do not arise when performing actual calculations because the physics makes us work in a corner of the structure where things are reasonable.
+The issue is that when one tries to prove theorems or general results on quantum theory, one currently is forced to consider the entire Hilbert space.
+It is unreasonable to expect that all theorems valid on all physical cases would have valid generalizations on unphysical ones, and therefore we are likely missing useful results.
+We therefore wonder whether many of the technical issues we have in quantum theory (e.g.
+non existence of a valid measure for path integrals, inequivalent representation of interacting theories, non existence of the categorial tensor product for Hilbert spaces, having to deal with infinities and infinite correlations)\footnote{TODO: cite} are simply a consequence of the use, at a fundamental level, of unphysical objects.
 
@@
 Suppose the reader agrees with our finding, the subsequent question is: what is the alternative?
-If we are only interested in observable with finitely many possible outcomes, we can use finite dimensional spaces.
-But as soon we have discrete quantities with arbitrarily large values, such as the number of particles, or quantities that are continuous, like position and momentum, we need infinite dimensional spaces.
-If we allow infinite dimensional spaces, we need some way of controlling what infinite linear combinations are allowed or not allowed.
-Cauchy completeness is one such way, but if that is not physical, what should we use?
+We do need infinite dimensional spaces to handle unbound quantities, such as number of particles, or continuous quantities, such as position.
+If we can't use Cauchy completeness, how do we characterize the converge of infinite sequences?
 
-We saw that the main problem is having expectation values that diverge, therefore the naive idea would be to require that the expectation value for all possible variables converge.
-This would require the expectation for $e^x$, $e^{e^x}$, $e^{e^{e^x}}$ and so on to be finite as well, which seems rather extreme.
-But what is a reasonable point to stop?
-If the idea is that we need to be able to reconstruct the state from expectation values, we should note that bounded distributions can be recovered from all the moments of the distribution, from the expectation of $x^n$ for every finite $n$.
-It seems reasonable, then, to require the expectation of all polynomial of position and momentum to be finite.
-
-In quantum mechanics, this means requiring all expectations of $X^nP^m = -\imath \hbar x^n\partial_x ^m$ to be finite.
+A solution could be simply focusing on the expectation values of observables, and requiring a sufficient number of them to be finite.
+For example, we could require the expectation of all polynomial of position and momentum to be finite.
+In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m - P^mX^n)$ to be finite.
 This restricts our space to those functions that are infinitely smooth (i.e.
-all derivative exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions.
-Requiring finite expectation for all polynomials of position and momentum, then, does not give us the usual $L^2$ Hilbert space, it gives us the Schwartz space.
-Is this a better candidate?
+all derivative exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.
+Is the Schwartz space a better candidate?
 
@@
 Moreover, it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with an arbitrary level of precision.
-Or, equivalently, that once a map is defined over the Schwartz space, the map is uniquely extended over the Hilbert space.
 This is a good hint that we have thrown away only unphysical elements.
@@
 Limits in terms of Cauchy sequences are not guaranteed, but limits in terms of expectation values are guaranteed: if we have a sequence of states such that the limit of the expectation converge, then we are guaranteed that the limit is also a Schwartz function.
+Schwartz functions, in fact, are uniquely identified by the expectation of all polynomial of position and momentum, effectively guaranteed that they can be distinguished by quantum tomography.
 
@@
 That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property.
-This is not, in retrospect, a nice property that would be convenient for physicists.
 The theory of tempered distribution, which allows us to define the Dirac $\delta$-function, is mathematically based, at its core, over the Schwartz space.
@@
 
-On the downside, many of the tools we use indeed assume a Hilbert space and many theorems would not hold without Cauchy completeness.
-There are two responses to this issue.
-First of all, there is no reason to prohibit that calculation are performed on the larger, unphysical space.
-When solving an integral, we do sometimes extend the domain for real to complex values, which are unphysical, because complex analysis has nicer mathematical features.
-The result so found, however, is still the valid result on the real domain.
-Similarly, we can pose the problem on the Schwartz space, extend to the Hilbert space for calculation, and then bring the result back to the Schwart space.
-However, there is a second way to look at this issue.
-Given that the focus has been squarely on Hilbert spaces, we do not know which mathematical results will hold, or even whether other results may be found.
-This is not something that the authors of this paper can answer and neither the mathematicians that were asked.
-This requires a serious research effort on the mathematical underpinning of multiple areas, which brings us to our closing point.
-
-This effort is never going to happen if the physics community, as whole, do not generate enough interest on the subject.
-There may be a better series of mathematical tools and results that stem from a more precise specification of our physical requirements in terms of mathematical definitions.
-It is unlikely that the mathematical world will give us these tools unprompted.
-It is the job of us physicists to understand exactly what are ramifications of the definitions in terms of what physical objects would be left out or what unphysical objects would be brought in.
+While determining whether Schwartz spaces are the correct space to capture quantum states goes beyond the scope of this article, we do want to show that alternatives exist.
+Exploring these alternatives fully, to see exactly what theorems still hold which do not, and whether others are possible, is a significant effort that is extremely important and long overdue.
 
@@
 
-% TODO: hint that the unitary transformations on Schwartz space are a subset of the ones on the Hilbert, and therefore they have unique inverse; operators we are interested in are uniquely extended to the Hilbert space.
+We have made the case that Hilbert spaces in quantum mechanics are \textbf{unphysical} as they necessarily bring in states with infinite expectation values, coordinate transformations that turn finite expectations in infinite ones, and unitary transformations that do the same in finite time.
+Theorems that are valid on Hilbert spaces must be valid over these unphysical cases, meaning that we are possibly missing results and techniques that would be physically interesting.
+
+The more general problem is that the physics community has become complacent in simply accepting mathematical definitions, without properly understanding their limit of applicability for physical theories.
+We believe that meaningful progress on the foundations of physics cannot happen without understanding the implicit assumptions embedded in the most basic mathematical tools.
 
@@
 
+\section{TODO}
+
+of a We saw that the main problem is having expectation values that diverge,
+
+If we are only interested in observable with finitely many possible outcomes, we can use finite dimensional spaces.
+But as soon we have discrete quantities with arbitrarily large values, such as the number of particles, or quantities that are continuous, like position and momentum, we need infinite dimensional spaces.
+If we allow infinite dimensional spaces, we need some way of controlling what infinite linear combinations are allowed or not allowed.
+Cauchy completeness is one such way, but if that is not physical, what should we use?
+
+We saw that the main problem is having expectation values that diverge, therefore the naive idea would be to require that the expectation value for all possible variables converge.
+This would require the expectation for $e^x$, $e^{e^x}$, $e^{e^{e^x}}$ and so on to be finite as well, which seems rather extreme.
+But what is a reasonable point to stop?
+If the idea is that we need to be able to reconstruct the state from expectation values, we should note that bounded distributions can be recovered from all the moments of the distribution, from the expectation of $x^n$ for every finite $n$.
+It seems reasonable, then, to require the expectation of all polynomial of position and momentum to be finite.
+
+In quantum mechanics, this means requiring all expectations of $X^nP^m = -\imath \hbar x^n\partial_x ^m$ to be finite.
+This restricts our space to those functions that are infinitely smooth (i.e.
+all derivative exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions.
+Requiring finite expectation for all polynomials of position and momentum, then, does not give us the usual $L^2$ Hilbert space, it gives us the Schwartz space.
+Is this a better candidate?
+
+First of all, the Schwartz space is a subspace of the $L^2$ Hilbert space taken as an inner product space.
+That is, all Schwartz functions are elements of the Hilbert space, they are closed under linear combinations and have the same inner product.
+It is not, however, complete, which is what we want.
+Moreover, it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with an arbitrary level of precision.
+Or, equivalently, that once a map is defined over the Schwartz space, the map is uniquely extended over the Hilbert space.
+This is a good hint that we have thrown away only unphysical elements.
+
+If we restrict ourselves to the Schwartz space, then, the only unitary transformations we are interested in are those that map Schwartz function to Schwartz functions.
+The change of coordinate we defined in the previous section, then, is ruled out.
+Limits in terms of Cauchy sequences are not guaranteed, but limits in terms of expectation values are guaranteed: if we have a sequence of states such that the limit of the expectation converge, then we are guaranteed that the limit is also a Schwartz function.
+
+The Schwartz space has also a very important property: it is the only space that is closed under the Fourier transform.
+That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property.
+This is not, in retrospect, a nice property that would be convenient for physicists.
+The theory of tempered distribution, which allows us to define the Dirac $\delta$-function, is mathematically based, at its core, over the Schwartz space.
+Therefore, this space already plays a fundamental role in the definitions of the tools we already use.
+
+On the downside, many of the tools we use indeed assume a Hilbert space and many theorems would not hold without Cauchy completeness.
+There are two responses to this issue.
+First of all, there is no reason to prohibit that calculation are performed on the larger, unphysical space.
+When solving an integral, we do sometimes extend the domain for real to complex values, which are unphysical, because complex analysis has nicer mathematical features.
+The result so found, however, is still the valid result on the real domain.
+Similarly, we can pose the problem on the Schwartz space, extend to the Hilbert space for calculation, and then bring the result back to the Schwart space.
+However, there is a second way to look at this issue.
+Given that the focus has been squarely on Hilbert spaces, we do not know which mathematical results will hold, or even whether other results may be found.
+This is not something that the authors of this paper can answer and neither the mathematicians that were asked.
+This requires a serious research effort on the mathematical underpinning of multiple areas, which brings us to our closing point.
+
+This effort is never going to happen if the physics community, as whole, do not generate enough interest on the subject.
+There may be a better series of mathematical tools and results that stem from a more precise specification of our physical requirements in terms of mathematical definitions.
+It is unlikely that the mathematical world will give us these tools unprompted.
+It is the job of us physicists to understand exactly what are ramifications of the definitions in terms of what physical objects would be left out or what unphysical objects would be brought in.
+
+% TODO: hint that the unitary transformations on Schwartz space are a subset of the ones on the Hilbert, and therefore they have unique inverse; operators we are interested in are uniquely extended to the Hilbert space.
+
 \end{document}

### 2023-06-21 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Cleaning up the appendix

@@
 
-We give the full mathematical detail in support of the argument laid out in the main paper
-
-A vector space is a set where the elements can be summed and multiplied by a constant.
-In quantum mechanics, the set represents pure states and the linear combination represent superposition of states.
-One may already accept this as physical, as superpositions can be created in a lab when, for example, a particle goes through a beam splitter.
-However, there are two difficulties.
-
-First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs.
-This is an area where different interpretations may differ, and we do not want our physical justification to depend on our interpretation.
-We want to decide whether the mathematical representation is physical regardless of the interpretation.
-Second, the actual states are not represented by vectors, but by rays.
-The superposition, then, is in terms of objects that are not themselves strictly physical.
-For these two reasons, we want to find another way an equivalent characterization of the linearity of the space that is fully physical.
-
-Let us consider the space of density operators, which can be used to characterize both pure and mixed state.
-These objects are invariant under change of phase and are in one-to-one correspondence with the rays of the space, not with the vectors.
-The density operators have a linear structure that represent statistical mixtures.
-This is a necessary structure because preparing statistical mixtures is something we can always do physically, and therefore something a physically meaningful mathematical definition must provide.
-However, the linear structure of density operators is not the same as the linear structure of the Hilbert space.
-
-When we write
-$$c_1|\psi_1\> + c_2 |\psi_2\>$$
-and 
-$$p_1|\psi_1\>\<\psi_1| + p_2|\psi_2\>\<\psi_2|$$
-the meaning of the $+$ symbols is different.
-The first operations is a complex linear combination of two pure states that returns another pure state, while the second is a real linear combination of two pure states that returns a mixed state.
-However, the two are so intertwined that one can be recovered by the other.
-
-As we show in the appendix, a pure state $\psi$ is a superposition of two other pure states $\phi_1$ and $\phi_2$ if and only if there exists another pure states $\bar{\psi}$ such that we can mix $\psi$ and $\bar{\psi}$ to obtain a mixed state that can also be obtained by mixing $\phi_1$ and $\phi_2$.
-Geometrically, consider a Bloch sphere.
-If $\psi$ is a superposition of $\phi_1$ and $\phi_2$ then all three states are on the surface of the same sphere.
-The chord connecting $\phi_1$ and $\phi_2$ represents the space of all possible mixtures of $\phi_1$ and $\phi_2$.
-Now, take a point $\rho$ on that chord, and consider the line that passes through $\rho$ and $\psi$.
-The same line will traverse the sphere at another point, which we call $\bar{\psi}$.
-Since $\rho$ is on the chord that connects $\psi$ and  $\bar{\psi}$, it is a mixture of the two.
-Therefore the vector space structure can be recovered by knowing how pure state mix.
-
-Note that the fact that the same statistical mixture can be decomposes in multiple ways is a characteristic feature of quantum mechanics, that is not present in classical mechanics.
-Using the same characterization, then, we consistently conclude that in classical mechanics there are no superpositions precisely because, given a pure state $\psi$, we can never find a corresponding state $\bar{\psi}$.
-
-The argument, then, goes like this.
-On physical grounds, we must be able to construct mixture.
-On physical grounds, we must be able to tell whether the same mixture can be prepared in different ways.
-The vector space structure captures exactly which pairs of pure states produce indistinguishable mixtures.
-The vector space structure exactly captures that structure and therefore it is fully physically justified.
-\textbf{The vector space structure is physical.}
-
-\subsection{Inner product}
-
-The inner product defines a geometrical structure on top of the vector space.
-Namely, it gives an idea of size of each vector and of angle between vectors.
-In quantum mechanics, the square of the inner product
-\begin{equation}
-	p(\psi_1, \psi_2) = |\<\psi_1| \psi_2\>|^2.
-\end{equation}
-represent the probability of measuring one state having prepared another.
-One may already accept this as physical, as these probabilities define all observable features of quantum mechanics.
-There is, again, a difficulty.
-
-Understanding the inner product as giving the probability of measurements also brings in one of the most controversial aspects of the theory: measurements.
-This is, again, a problem since we want our reasoning to be independent of interpretation.
-Another issue is that, technically, the inner product is linked to decomposition in linear combinations, to superposition, which are done at equal time.
-The probability above, instead, is a probability of transition.
-
-Given that we have characterized superposition in terms of mixtures, we should see whether the inner product characterizes something fundamental about them, and indeed it does.
-Given two pure states, the entropy of the mixture is a strictly monotonic, and therefore invertible, function of the square of the inner product.
-The greatest entropy is given by mixing orthogonal states, which gradually decreases as the angle in the inner product space decreases.
-We can then understand the geometrical structure of the pure quantum states as an entropic structure, much like it happens for classical mechanics.\footnote{In classical mechanics, the Liouville measure returns the count of states within a region, which is used in the fundamental postulate of statistical mechanics.
-The symplectic form gives us the Liouville measure for each independent degree of freedom.}
-
-The argument, then, goes like this.
-On physical grounds, for each statistical mixture, the thermodynamic entropy must be defined.
-Therefore it must be defined for equal mixtures of two pure states as well.
-The inner product exactly captures how that quantity combines and therefore it is fully physically justified.
-\textbf{The inner product structure is physical.}
+In this appendix we provide the same arguments presented in the main body of the article expanded with the full mathematical detail.
 
@@
 
-First we want to show that the vector space structure is not, by itself, physical.
+The first task is to show that the vector space structure is not, by itself, physical.
 Since quantum states are represented by rays of a Hilbert space, and since rays can be defined simply on top of the vector space structure alone, one may think that the vector space structure, by itself, is physical.
@@
 
-The following proposition tells us that there are infinitely many ways to decompose a vector representing a state in terms of vectors that represent two other states.
+To show the problem, we prove the following proposition that tells us that there are infinitely many ways to decompose a vector representing a state in terms of vectors that represent two other states, meaning that the coefficients of the linear combination are arbitrary.
 \begin{prop}\label{vector_insufficient}[The vector space structure is not enough to uniquely identify superpositions.] Let $\mathcal{H}$ be a complex vector space and let $\textbf{P}(\mathcal{H})$ be its projective space.
@@
 The intuition that a state has a single linear decomposition in terms of other vectors requires, first of all, that the vectors are normalized.
-To identify the coefficient as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires the decomposition to be done on orthogonal vectors.\footnote{TODO: mention Gram-Schmidt orthonormalization} That is, linear combinations require the inner product to be understood physically as superpositions.
+To identify the coefficient as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires the decomposition to be done on orthogonal vectors.\footnote{Mathematically, this means turning the vectors $\{\phi_i\}$ into an orthonormal set using Gram-Schmidt orthonormalization, which requires the inner product.} That is, linear combinations require the inner product to be understood physically as superpositions.
 The vector space structure by itself is therefore insufficient to provide a clear connection to physics.
@@
 First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs.
-This is an area where different interpretations may differ,\footnote{TODO: add citations} some arguing that they are just mathematical representations, and the physicality of a mathematical representation should not be interpretation dependent.
+This is an area where different interpretations may differ,\footnote{TODO: add citations} and the physicality of a mathematical representation should not be interpretation dependent.
 Second, the Born rule, in its standard characterization, brings in one of the most controversial aspects of the theory, measurements, which suffers from an interpretation problem as well.
@@
 The idea is to reformulate the requirements that lead to superposition and the Born rules in terms of equivalent requirements of statistical mixtures.
-Mathematically, these will be in terms of the space of density operators in stead of the vector space itself.\footnote{Similar efforts can be found in the literature of Quantum Logic and of Generalized Probability Theories.
-TODO: add citations}.
+Mathematically, these will be in terms of the space of density operators instead of the vector space itself.\footnote{Similar techniques can be found in the literature of Quantum Logic and of Generalized Probability Theories.
+TODO: add citations}
 
@@
 For example, if we take the maximally mixed state for a spin 1/2 system, this can be implemented with an equal mixture of spin up and spin down, or with an equal mixture of spin left and spin right.
-The idea is that a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ exactly if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
+The idea is that a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ exactly because there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
+Hilbert space subspaces are grouping together pure states that can provide the same statistical mixtures.
 To keep the reasoning as general as possible, we will not require $\{\phi_i\}$ to be orthogonal to each other.
@@
 Let $\{\psi_i\}, \{\phi_j\} \in \mathcal{H}$ be two finite sets of normalized vectors.
-Suppose there exists a density operator that can be expressed as a strict convex combination of either set, that is we can find two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}^+$ such that
+Suppose there exists a density operator that can be expressed as a strict convex combination of either set, that is we can find two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}_{>0}$ such that
 	\begin{enumerate}
@@
 \begin{proof}
-	The set of all possible density operators is a convex set, which can be understood as a subset of a vector space $V$.
+	The set of all possible density operators is a convex set, which can be characterized as a subset of a vector space $V$.
 Let $\rho$ be a density operator such that $\<\psi|\rho|\psi\>\neq 0$.
+This will be a point in the convex set.
 If $\rho = |\psi\>\<\psi|$, we are done.
@@
 If it is an extreme point, we are done as $\rho$ is an affine combination of $\psi$ and $\rho_1$.
-If not, the boundary of the convex set in a neighborhood of $rho_1$ is not strictly convex, and therefore is a subspace $V_1$ of the vector space $V$.
+If not, $\rho_1$ is on a part of the boundary that is not strictly convex.
+That region will be a convex set constrained on a subspace $V_1$ of $V$.
 Note that $V_1$ does not include $\psi$ and it does not span the direction identified by $\rho_1$ and $\psi$.
-Furthermore, we can't have $\rho_1 = \rho$ as $\<\psi|\rho|\psi\>\neq 0$.
+This means that we can't have $\rho_1 = \rho$ as $\<\psi|\rho|\psi\>\neq 0$.
 Therefore we have that $\rho$ is an affine combination of $\psi$ and $\rho_1$, which must be a density operator of finite rank lower than $\rho$.
@@
 Then $\psi$ is a superposition of $\{\phi_j\}$ if and only if there exists a density operator $\rho$ that can be expressed both as a strict convex combination of $\{\phi_i\}$ and of another set of normalized vectors that contain $\psi$.
-That is, we can find a countable set of normalized vectors $\{\psi_i\} \in \mathcal{H}$ and two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}^+$ such that
+That is, we can find a finite set of normalized vectors $\{\psi_i\} \in \mathcal{H}$ and two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}_{>0}$ such that
 \begin{enumerate}
@@
 \begin{proof}
-	Suppose $\psi$ is a normalized state that can be written as $\psi = \sum_{i=1}^n c_i \phi_i$ where $\{\phi_i\} \in \mathcal{H}$ a finte sets of normalized vectors that do not contain $\psi$.
-Since $\{\phi_j\}$ are not necessarily orthogonal, it is not necessarily true that $|\<\psi|\phi_i\>| = |c_i|^2 \neq 0$.
-However, since $\psi$ is normalized, there must be at least two vectors in $\{\phi_i\}$ such that $|\<\psi|\phi_i\>| \neq 0$.
+	Suppose $\psi$ is a normalized state that can be written as $\psi = \sum_{i=1}^n c_i \phi_i$ where $\{\phi_i\} \in \mathcal{H}$ is a finite sets of normalized vectors that do not contain $\psi$.
+Since $\{\phi_j\}$ are not necessarily orthogonal, it is not necessarily true that $|\<\psi|\phi_i\>| = |c_i|^2$ or that $|\<\psi|\phi_i\>| \neq 0$ for all $i$.
+However, since $\psi$ is normalized and $\psi \neq \phi_i$ for all $i$, there must be at least two vectors in $\{\phi_i\}$ such that $|\<\psi|\phi_i\>| \neq 0$.
 We can then find $\{p_i\} \in \mathbb{R}+$ for which $\sum_i p_i =1$, such that $\rho = \sum_i p_i |\phi_j\>\<\phi_j|$ and $\<\psi|\rho|\psi\> \neq 0$.
@@
 
-We can now show that the inner product space structure is physical in the sense that it is exactly the structure that is needed to model quantum mixed states and their entropy.
-One key difference between classical and quantum mechanics is that there is no single decomposition of quantum mixed states in terms of pure states.
-We can show that relationships of superposition are equivalent to relationships of multiple decomposition.
-
-\begin{proof}
-	Let's first address the trivial case in which $\phi_1 = \phi_2$.
-Suppose that $\psi = c_1 \phi_1 + c_2 \phi_2$, then $\psi = \phi_1$.
-We set $\bar{\psi} = \psi$ and find $p_1|\psi\>\<\psi| + p_2|\bar{\psi}\>\<\bar{\psi}| = \frac{1}{2}|\phi_1\>\<\phi_1| + \frac{1}{2}|\phi_2\>\<\phi_2|$ for any $p_1, p_2 \in \mathbb{R}$ such that $p_1+p_2=1$.
-Conversely, suppose $p_1|\psi\>\<\psi| + p_2|\bar{\psi}\>\<\bar{\psi}| = \frac{1}{2}|\phi_1\>\<\phi_1| + \frac{1}{2}|\phi_2\>\<\phi_2|$ for some $p_1, p_2 \in \mathbb{R}$ such that $p_1+p_2=1$.
-Then $p_1|\psi\>\<\psi| + p_2|\bar{\psi}\>\<\bar{\psi}| = |\phi_1\>\<\phi_1|$ which can be true only if $\psi=\bar{\psi}=\phi_1$.
-
-	Now let's address the case in which $\phi_1 \neq \phi_2$.
-Suppose that $\psi = c_1 \phi_1 + c_2 \phi_2$.
-Then $\psi$ is part of the subspace defined by $\phi_1$ and $\phi_2$, which means all the states can be represented as points on the surface of a Bloch sphere.
-The mixed state $\rho=\frac{1}{2}|\phi_1\>\<\phi_1| + \frac{1}{2}|\phi_2\>\<\phi_2|$ is the midpoint of the chord connecting $\phi_1$ and $\phi_2$.
-Now, take the line that connects $\rho$ with $\psi$.
-This line will traverse the sphere in another point $\hat{\psi}$.
-As $\rho$ sits on the chord between $\psi$ and $\bar{\psi}$, it is a mixture of the two states for some $p_1$ and $p_2$.
-Conversely, suppose $\rho = p_1|\psi\>\<\psi| + p_2|\bar{\psi}\>\<\bar{\psi}| = \frac{1}{2}|\phi_1\>\<\phi_1| + \frac{1}{2}|\phi_2\>\<\phi_2|$.
-This means that the subspace spanned by $\psi$ and $\bar{\psi}$ is the same as the one spanned by $\phi_1$ and $\phi_2$, which means that $\psi = c_1 \phi_1 + c_2 \phi_2$ for some $c_1, c_2 \in \mathbb{C}$ such that $|p_1|^2+|p_2|^2=1$.
-\end{proof}
-
-As for the inner product, we show that it can be fully characterized, instead of by the Born rule, but the entropy of equal mixtures over state pairs.
-This gives us a geometric understanding of the inner product at equal time, instead of a probability of transition.
+As for the inner product, we show that the Born rule can be understood as characterizing the entropy of equal mixtures over state pairs.
+This gives us a geometric understanding of the inner product at equal time, instead of a probability of transition, which brings it more in line with the use during superposition.
+It should also be noted that the von Neumann entropy of a mixture can be understood as the minimum Shannon entropy over all possible decompositions,\footnote{TODO: cite} which again shows the connection between the geometry of the convex space of statistical ensembles and the inner product.
 
@@
 
-We are then able to argue for the following
+Having shown the connection between superposition and non-unique decomposition of mixed state, and between the Born rule and the entropy of mixtures, we can now use these to justify the physicality of the complex inner product space structure for quantum states.
 \begin{prop}

### 2023-06-22 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Added appendix on unphysicality of Hilbert spaces

@@
 Mathematically, this means that the space includes the limit of every Cauchy sequence; physically, this property is more difficult to characterize.
-We can look at an equivalent characterization:\footnote{Roman 2008, p.
+We can look at an equivalent characterization:\cite{roman_2008}\footnote{Roman 2008, p.
 330 Theorem 13.8
@@
 \begin{prop}
-	The description of quantum states as rays of a complex inner product space is physically justified.
+	Complex inner product spaces are physical when used to represent the state space of quantum systems.
 \end{prop}
 \begin{justification}
-	Since physical preparations and measurements are most of the type not perfect, a physical theory must define statistical ensembles.
-Part of that definition includes keeping track statistical ensembles under statistical mixtures, that is which particular mixture is obtained by mixing particular states in particular proportions.
+	Since physical preparations and measurements are most of the time not perfect, a physical theory must define statistical ensembles.
+Part of that definition includes keeping track of statistical ensembles under statistical mixtures, that is which particular mixture is obtained by mixing particular states in particular proportions.
 A particular feature of quantum mixed states is that, unlike classical states, they do not in general provide a unique decomposition in terms of pure states.
@@
 
-\section{TODO}
+\section{The unphysicality of completeness}
 
-of a We saw that the main problem is having expectation values that diverge,
+To show that the completeness requirement on complex inner product spaces that represent quantum states is unphysical, we need to show that this definition forces the state space to include elements that cannot have physical correspondents.
+
+The Hellinger–Toeplitz theorem states that a self-adjoint operator $O$ defined on the whole Hilbert space $\mathcal{H}$ is necessarily bound.
+Therefore, an unbound operator $X$ cannot be defined on the whole space, meaning that the operator must diverge or be ill-defined on a subset of $\mathcal{H}$.
+To see that the problem is the completeness requirement of the Hilbert space, we prove the following:
+\begin{prop}
+	Let $\mathcal{H}$ be a Hilbert space and $X : D(X) \to Y$ a unbounded operator where $D(X)$ is the domain of $X$ and $Y$ is another Hilbert space.
+Then $D(X)$ is an inner product space that is not complete.
+\end{prop}
+\begin{proof}
+	The domain $D(X)$ of the operator is a subspace of $\mathcal{H}$.
+In fact, if $||X(v_1)||_Y$ and $||X(v_2)||_Y$ are finite, so will $||X(a_1 v_1+a_2v_2)||_Y$.
+Therefore $D(X)$ is an inner product space.
+It cannot be complete by the converse of the Hellinger–Toeplitz theorem.
+\end{proof}
+
+To understand the physical significance, consider the position operator $X : D(X) \to \mathcal{H}$.
+If we take a normalized vector $\psi \in \mathcal{H}$, the norm $||X\psi|| = \<X\psi|X\psi\> = E[X^2|\psi]$ is equal to the expectation of the position squared.
+The operator is unbounded because the position can be arbitrarily large.
+We can therefore construct a Cauchy sequence $\{\psi_i\}$ such that $\lim\limits_{i \to \infty}||X\psi|| = +\infty$, that is take a sequence of vectors for which the expectation of position is larger and larger.
+Because of completeness, the limit of the Cauchy sequence will be in the Hilbert space.
+However, because the expectation diverges, the limit cannot be in the domain of $X$.
+Cauchy completeness forces us to include ``states'' for which the position is ill-defined.
+
+The matter would still be manageable if we had a way to keep these degenerate cases isolated, but this is not the case. In general, a unitary operator can move vectors in and out of the domain.
+\begin{prop}
+	Let $\mathcal{H}$ be a Hilbert space, $X : D(X) \to Y$ a unbounded operator and $U : \mathcal{H} \to \mathcal{H}$ a unitary operator. Then, in general, $U(D(X)) \notin D(X)$.
+\end{prop}
+\begin{proof}
+	Let $\psi, \phi \in \mathcal{H}$ be two normalized vectors.
+Then we can construct a unitary operator $U$ such that $\phi = U(\psi)$.
+For example, we can take the operator that performs a rotation in the subspace identified by $\psi$ and $\phi$ and leaves the orthogonal subspace unchanged.
+We can now take $\psi$ to be in $D(X)$ while $\phi$ to not be in $D(X)$.
+Therefore, in general, a unitary operator can move vectors in and out the domain of an unbounded operator.
+\end{proof}
+
+To understand the physical significance, note that unitary operators are isomorphisms between Hilbert spaces.
+That is, mathematically, two Hilbert spaces are considered indistinguishable if there exists a unitary transformation between them.
+As we saw, unitary operations can change the domain in which unbound operators are well defined: they can map states for which position is well defined onto states for which it is not.
+While the mathematical structure remains unchanged, the physical significance hasn't.
+
+More concretely, a change of variable $X$ to $Y$ would be a unitary operator as is just a change of representation.
+However, we can construct changes of variables such that the domain of $X$ is not exactly mapped to the domain of $Y$.
+Let $y=y(x)$ be our change of variable such that $y(0) = 0$.
+If $\psi(x)$ is a wavefunction expressed over $x$, and $\phi(y)$ represents the same state as a wavefunction over $y$, we have the following relationship:
+\begin{equation}
+		\int_{0}^{y(x)} \phi^\dagger(y) \phi(y) dy = \int_{0}^{x} \psi^\dagger(x) \psi(x) dx
+\end{equation}
+To find a coordinate transformation that take finite expectations to infinite expectation, we set the initial and final wavefunctions as follow:
+\begin{align}
+	\psi(x) &= \sqrt{\frac{e^{-x^2}}{\sqrt{\pi}}} \\
+	\phi(y) &= \frac{1}{\sqrt{\pi(y^2 + 1)}}.
+\end{align}
+The first distribution is a Gaussian, and therefore all moments of the distribution, the expectation for $x^n$, are finite.
+The second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all moments above the first.
+We find
+\begin{equation}
+	\begin{aligned}
+		\int_{0}^{y(x)} \frac{1}{\pi(y^2 + 1)} dy &= \int_{0}^{x} \frac{e^{-x^2}}{\sqrt{\pi}} dx \\
+		\frac{\tan^{-1}(y(x))}{\pi} &= \frac{\erf(x)}{2} \\
+		y(x) &= \tan \left(\frac{\pi}{2}\erf(x)\right). \\
+	\end{aligned}
+\end{equation}
+The graph of the above function increase so fast that, when plotted, it looks like it has a vertical asymptote.
+Nonetheless, it is an invertible function over the whole domain, and will induce a unitary transformation on the Hilbert space.\footnote{Note that we could implement a similar coordinate transformations in classical mechanics, which would also change finite moments of a distribution into infinite ones.
+It may be that, like differentiable manifolds are constrained to only differentiable coordinate transformation, we need ``statistical manifolds'' that are constrained to coordinate transformations that preserve the finiteness of moments of distributions.
+This is a topic that requires further analysis.}
+
+Unitary operators are also used in quantum theory to represent time evolution.
+Typically, we have a family of unitary operators $\{U_t\}_{t \in \mathbb{R}}$ that is strongly continuous (i.e.
+$\lim_{t \to t_0} U_t = U_{t_0}$) and respect composition in time (i.e.
+$U_t U_s = U_t+s$).
+We can turn the above transformation into an evolution by setting
+\begin{equation}
+	y(x, t) = \cos(\omega t) x + \sin(\omega t) \tan \left(\frac{\pi}{2}\erf(x)\right).
+\end{equation}
+The idea is that the evolution will simply stretch and shrink the wavefunction intermittently, potentially oscillating expectation values from finite to infinite and viceversa in finite time.
+Clearly, this is not a meaningful time evolution, yet it will correspond to a strongly continuous one-parameter unitary group and, by Stone's theorem, will even admit a Hamiltonian.
+
+The mathematician that points out that this all known, and there is nothing new, is missing the point.
+The physicists that notes that, yes, those elements are clearly unphysical, but I don't use them, so I do not care, is also missing the point.
+The point is that, physically, we will want to make general statements such as ``for all observer X is valid'' or ``under all evolution Y is valid''.
+Given that Hilbert spaces mix physically clearly meaningful object with clearly unphysical ones, we have no way to make those statements precisely.
+
+We can conclude the discussion with the following
+\begin{prop}
+	Infinite dimensional complex Hilbert spaces are unphysical when used to represent the state space of quantum systems due to completeness.
+\end{prop}
+\begin{justification}
+	If a quantum system is modeled by a finite dimensional space, and since every finite dimensional space is complete, finite dimensional complex Hilbert spaces are physical.
+
+	Some physical system that are described by unbounded discrete quantities (e.g.
+number of particles) or by continuous quantities (e.g.
+position) which require infinite dimensional spaces.
+Infinite dimensional complex inner product spaces, then, are needed to represent quantum system.
+
+	An unbounded operator cannot be defined on all the elements of a Hilbert space by the converse of the Hellinger-Toeplitz theorem.
+This means that if represents states as rays on a Hilbert space and observables as operators, the unbounded observable will not be well defined, not even as a statistical quantity, on all states.
+In particular, whether the statistical properties of an observable are well defined, whether it is possible to experimentally identify a state through tomography, would be observer dependent.
+Moreover, whether the expectation of an observer is finite or not would be something that can change in finite time.
+
+	On physical ground, we must have that an observable has well defined statistical properties over all states, that this property is covariant and that is preserved by time evolution.
+Hilbert spaces do not satisfy this requirement and are therefore unphysical.
+\end{justification}
+
+\section{The possible physicality of Schwartz spaces}
+
+Recall that an inner product space $V$ is complete if every Cauchy sequence in $V$ converges in $V$.
+This definition is difficult to characterize physically, therefore we use an equivalent condition: an inner product space $V$ is complete is complete if and only if all absolutely convergent series converge in $V$ (see theorem 13.8 in \cite{roman_2008}).
+This reframes the requirement in terms of convergence of linear combinations, and therefore superpositions.
+
+We now show that if we have an unbounded operator, like position, completeness requires the existence of pure states for which
 
@@
 
+\section{A final remark on the mathematical foundations of physical theories}
+
+When mathematicians and philosophers noticed that mathematical theories at the time allowed to express inconsistencies (e.g.
+Russell's paradox, Berry paradox, ...), effort was put into strengthening the foundations of mathematics so that such paradoxes were either resolved or couldn't be expressed.
+With regard to the mathematical foundations of physical theories, we have a similar problem: we can write statements that, though mathematically consistent, are physically inconsistent.
+Looking for these problems and addressing them is ultimately what will lead to a solid foundation, which, we believe, is the prerequisite to solve other outstanding issues in fundamental physics.
+
 \end{document}

### 2023-06-23 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Adding appendix on Schwartz spaces

@@
 
-Recall that an inner product space $V$ is complete if every Cauchy sequence in $V$ converges in $V$.
-This definition is difficult to characterize physically, therefore we use an equivalent condition: an inner product space $V$ is complete is complete if and only if all absolutely convergent series converge in $V$ (see theorem 13.8 in \cite{roman_2008}).
-This reframes the requirement in terms of convergence of linear combinations, and therefore superpositions.
+We have found that:
+\begin{enumerate}
+	\item complex inner product spaces are needed
+	\item infinite dimensional spaces are needed
+	\item completeness is unphysical.
+\end{enumerate}
+In other words, we need a stricter way to characterize convergence within the state space so that we can manage the infinite dimensionality.
+Conceptually, we need to treat the infinities coming from unbound quantities as potential infinities, not actual infinities: we never actually have infinitely many particles or a system positioned at infinity, we have an infinite range of possible values.
 
-We now show that if we have an unbounded operator, like position, completeness requires the existence of pure states for which
+A naive idea would be to require that the expectation value for all possible variables converge.
+Unfortunately, this doesn't work.
+Suppose we have a distribution $\rho(x)$ over position $x$.
+We would be requiring the expectation of any function to be finite, which would include $e^x$, $e^{e^x}$, $e^{e^{e^x}}$ and so on.
+The only way this can work is if the function has compact support.
+Physically, this seems too restrictive as it excludes functions like normal distributions.
+At a closer investigation, the Paley-Wiener theorem tells us that the Fourier transform of a function with compact support is an entire function.
+The set of zeros of an entire function is either discrete or the full space.
+Therefore, we cannot have compact support in both position and momentum.
 
-If we are only interested in observable with finitely many possible outcomes, we can use finite dimensional spaces.
-But as soon we have discrete quantities with arbitrarily large values, such as the number of particles, or quantities that are continuous, like position and momentum, we need infinite dimensional spaces.
-If we allow infinite dimensional spaces, we need some way of controlling what infinite linear combinations are allowed or not allowed.
-Cauchy completeness is one such way, but if that is not physical, what should we use?
-
-We saw that the main problem is having expectation values that diverge, therefore the naive idea would be to require that the expectation value for all possible variables converge.
-This would require the expectation for $e^x$, $e^{e^x}$, $e^{e^{e^x}}$ and so on to be finite as well, which seems rather extreme.
-But what is a reasonable point to stop?
-If the idea is that we need to be able to reconstruct the state from expectation values, we should note that bounded distributions can be recovered from all the moments of the distribution, from the expectation of $x^n$ for every finite $n$.
-It seems reasonable, then, to require the expectation of all polynomial of position and momentum to be finite.
-
+The only way out, then, is to require only a certain class of observables to have finite expectation.
+Physically, it would make sense to require exactly those observables that are needed to fully reconstruct the state through quantum tomography.
+A good set in that respect seems to be the set of polynomial of conjugate quantities, like position and momentum.
 In quantum mechanics, this means requiring all expectations of $X^nP^m = -\imath \hbar x^n\partial_x ^m$ to be finite.
@@
 Requiring finite expectation for all polynomials of position and momentum, then, does not give us the usual $L^2$ Hilbert space, it gives us the Schwartz space.
-Is this a better candidate?
 
-First of all, the Schwartz space is a subspace of the $L^2$ Hilbert space taken as an inner product space.
+Not only the expectations of all polynomials exist, but they are enough to identify the elements and to define limits.
+The Schwartz space is in fact a Fr\'echet, meaning that is a Hausdorff space, with the topology induced by the seminorms given by the expectations and it is complete with respect to those seminorms.
+Physically, this means that measurements on the polynomials (i.e.
+the topology of the seminorms) are enough to identify the states (i.e.
+the singletons, which are closed sets as per Hausdorff).
+Moreover, if we have a sequence for which all expectations converge, we are guaranteed to converge to another state (i.e.
+completeness with respect to the seminorms).
+If any expectation diverges, the limit is not in the space.
+This criteria of convergence seems a lot more physically sound.
+
+To compare Schwartz and Hilbert spaces, the Schwartz space is a subspace of the $L^2$ Hilbert space taken as an inner product space.
 That is, all Schwartz functions are elements of the Hilbert space, they are closed under linear combinations and have the same inner product.
-It is not, however, complete, which is what we want.
-Moreover, it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with an arbitrary level of precision.
+It is not complete and it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with arbitrary level of precision.
 Or, equivalently, that once a map is defined over the Schwartz space, the map is uniquely extended over the Hilbert space.
-This is a good hint that we have thrown away only unphysical elements.
+It would seem, then, that all we are losing is the behavior at  actual infinity.
+
+The Schwartz space has also a very important property: it is the only space that is closed under Fourier transform.
+That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property.
+Given the importance of the Fourier transform in quantum theory, this feels like an important property to have.
+Mathematically, the theory of tempered distribution, which allows us to define objects like the Dirac $\delta$-function, is built on top of the Schwartz space.
+Therefore, this space already plays a fundamental role in the definitions of the tools we already use.
 
 If we restrict ourselves to the Schwartz space, then, the only unitary transformations we are interested in are those that map Schwartz function to Schwartz functions.
-The change of coordinate we defined in the previous section, then, is ruled out.
-Limits in terms of Cauchy sequences are not guaranteed, but limits in terms of expectation values are guaranteed: if we have a sequence of states such that the limit of the expectation converge, then we are guaranteed that the limit is also a Schwartz function.
+These are special cases of unitary transformation over Hilbert spaces, and inherit all their properties, such as the existence of the inverse.
+The change of coordinate we defined in the previous section, then, is automatically ruled out.
+It is interesting to note that two Hilbert spaces connected by such unitary transformations have essentially have two separate theories of distributions as they start from two different sets of objects.
+In our mind, this shows exactly how two inequivalent physical representations can be mathematically equivalent.
+Our suspicion is that the problem of inequivalent mathematical representations in QFT is a consequence of adding another non-physical infinity.
 
-The Schwartz space has also a very important property: it is the only space that is closed under the Fourier transform.
-That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property.
-This is not, in retrospect, a nice property that would be convenient for physicists.
-The theory of tempered distribution, which allows us to define the Dirac $\delta$-function, is mathematically based, at its core, over the Schwartz space.
-Therefore, this space already plays a fundamental role in the definitions of the tools we already use.
+Though we believe the arguments to be thoroughly convincing, we are stopping short from declaring Schwartz spaces physical.
+First, because this is not the goal of the paper, which is just to show the unphysicality of Hilbert spaces.
+Second, because we would need to show the importance of expectation values of polynomials from experimental considerations.
+It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.
 
-On the downside, many of the tools we use indeed assume a Hilbert space and many theorems would not hold without Cauchy completeness.
-There are two responses to this issue.
-First of all, there is no reason to prohibit that calculation are performed on the larger, unphysical space.
+Now, one may be concerned that if we abandon Hilbert spaces we would also have to abandon many convenient tools built on top.
+This is not the case.
 When solving an integral, we do sometimes extend the domain for real to complex values, which are unphysical, because complex analysis has nicer mathematical features.
 The result so found, however, is still the valid result on the real domain.
-Similarly, we can pose the problem on the Schwartz space, extend to the Hilbert space for calculation, and then bring the result back to the Schwart space.
-However, there is a second way to look at this issue.
-Given that the focus has been squarely on Hilbert spaces, we do not know which mathematical results will hold, or even whether other results may be found.
-This is not something that the authors of this paper can answer and neither the mathematicians that were asked.
-This requires a serious research effort on the mathematical underpinning of multiple areas, which brings us to our closing point.
-
-This effort is never going to happen if the physics community, as whole, do not generate enough interest on the subject.
-There may be a better series of mathematical tools and results that stem from a more precise specification of our physical requirements in terms of mathematical definitions.
-It is unlikely that the mathematical world will give us these tools unprompted.
-It is the job of us physicists to understand exactly what are ramifications of the definitions in terms of what physical objects would be left out or what unphysical objects would be brought in.
-
-% TODO: hint that the unitary transformations on Schwartz space are a subset of the ones on the Hilbert, and therefore they have unique inverse; operators we are interested in are uniquely extended to the Hilbert space.
+Similarly, we can pose the problem on the Schwartz space, extend to the Hilbert space or to the space of distributions for calculation, and then bring the result back to the Schwartz space.
+What really need to be rethought, instead, is the construction of state space, especially under infinite operations.
+If we want to construct the state space for a composite system of potentially infinitely many subsystems, understanding exactly what infinities are physical and which aren't is exactly the issue.
+Therefore taking the Hilbert space, which we saw doesn't deal with physical infinities correctly, and making infinitely many copies of it is bound to create problems.
 

### 2023-06-23 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Another pass at the Schwartz section

@@
 
-Suppose the reader agrees with our finding, the subsequent question is: what is the alternative?
-We do need infinite dimensional spaces to handle unbound quantities, such as number of particles, or continuous quantities, such as position.
-If we can't use Cauchy completeness, how do we characterize the converge of infinite sequences?
+While we do need infinite dimensional spaces to handle unbound quantities, such as number of particles, or continuous quantities, such as position, it seems that Cauchy completeness is not the proper characterization.
+While finding the proper one goes beyond the scope of this work, we do want to show that alternatives are possible.
 
-A solution could be simply focusing on the expectation values of observables, and requiring a sufficient number of them to be finite.
-For example, we could require the expectation of all polynomial of position and momentum to be finite.
+A reasonable position would be to require the expectation of all polynomial of position and momentum to be finite.
 In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m - P^mX^n)$ to be finite.
@@
 all derivative exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.
-Is the Schwartz space a better candidate?
 
-First of all, the Schwartz space is a subspace of the $L^2$ Hilbert space taken as an inner product space.
-That is, all Schwartz functions are elements of the Hilbert space, they are closed under linear combinations and have the same inner product.
-It is not, however, complete, which is what we want.
-Moreover, it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with an arbitrary level of precision.
-This is a good hint that we have thrown away only unphysical elements.
-
-If we restrict ourselves to the Schwartz space, then, the only unitary transformations we are interested in are those that map Schwartz function to Schwartz functions.
-The change of coordinate we defined in the previous section, then, is ruled out.
-Limits in terms of Cauchy sequences are not guaranteed, but limits in terms of expectation values are guaranteed: if we have a sequence of states such that the limit of the expectation converge, then we are guaranteed that the limit is also a Schwartz function.
-Schwartz functions, in fact, are uniquely identified by the expectation of all polynomial of position and momentum, effectively guaranteed that they can be distinguished by quantum tomography.
-
-The Schwartz space has also a very important property: it is the only space that is closed under the Fourier transform.
-That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property.
-The theory of tempered distribution, which allows us to define the Dirac $\delta$-function, is mathematically based, at its core, over the Schwartz space.
-Therefore, this space already plays a fundamental role in the definitions of the tools we already use.
-
-While determining whether Schwartz spaces are the correct space to capture quantum states goes beyond the scope of this article, we do want to show that alternatives exist.
-Exploring these alternatives fully, to see exactly what theorems still hold which do not, and whether others are possible, is a significant effort that is extremely important and long overdue.
+The Schwartz space has a number of other interesting properties, explored in a bit more detail in the appendix. For example:
+\begin{itemize}
+	\item it is fully identified by the expectations - two Schwartz are the same if and only if all expectations are the same; quantum tomography on all polynomials fully reconstructs the state
+	\item it is inner product space with the same norm of $L^2$
+	\item it is a dense subset of $L^2$ Hilbert space - any element of the Hilbert space can be approximated by a Schwartz function with arbitrary level of precision
+	\item it is complete with respect to the expectations - a sequence of functions for which all expectations of all polynomials of position and momentum converge will converge in the Schwartz space  
+	\item its closure over all Cauchy sequences recovers the Hilbert space - the only thing missing are the Cauchy sequences for which the limit of at least one expectation is not well defined
+	\item it is the only space closed under Fourier transform - the Fourier transform of a Schwartz function is a Schwartz function
+	\item it plays a fundamental role in the theory of distributions - objects like the $\delta$-function are mathematically defined on top of the Schwartz space
+\end{itemize}
+These properties make the Schwartz space a much reasonable candidate to capture the physics.
+Convergence over vectors in an abstract space is replaced by converges of expectation values of actual observables.
+The point is that alternatives exist and should be fully explored.
 
@@
 Physically, this seems too restrictive as it excludes functions like normal distributions.
-At a closer investigation, the Paley-Wiener theorem tells us that the Fourier transform of a function with compact support is an entire function.
-The set of zeros of an entire function is either discrete or the full space.
-Therefore, we cannot have compact support in both position and momentum.
+At a closer investigation, the Paley-Wiener theorem\footnote{TODO: find citation} tells us that the Fourier transform of a function with compact support is an entire function, and entire function do not have compact support.\footnote{TODO: find citation} Therefore, we cannot have compact support in both position and momentum.
+\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
+Therefore the space of functions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}
 
@@
 the topology of the seminorms) are enough to identify the states (i.e.
-the singletons, which are closed sets as per Hausdorff).
-Moreover, if we have a sequence for which all expectations converge, we are guaranteed to converge to another state (i.e.
+the singletons, which are closed sets as per Hausdorff).\footnote{This is not true in classical mechanics.
+Even in the single variable case, there are distinct distributions that have the same moments.
+However, if we assume compact support, distinct distributions have distinct moments even in the multi variable case.\cite{moment_problem_2017} It is curious that the moment problem in quantum mechanics works differently.} Moreover, if we have a sequence for which all expectations converge, we are guaranteed to converge to another state (i.e.
 completeness with respect to the seminorms).
@@
 First, because this is not the goal of the paper, which is just to show the unphysicality of Hilbert spaces.
-Second, because we would need to show the importance of expectation values of polynomials from experimental considerations.
-It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.
+Second, because we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The fact that the moment problems works differently in classical and quantum mechanics may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.\footnote{One of the issue is the fact that uniqueness works}
 

### 2023-06-25 Articles/2023-HilbertQuantum/HilbertQuantum.tex
It looks like a lot of changes, but it's just rewriting sentences that were unclear, fixing typos, adding more references, and adding some TO DO

@@
 157]{browder_hilberts_1976}): to axiomatize the physical sciences.
-But, despite the undeniable importance of wanting to set physics on a firm mathematical footing, we must remember that by ``axioms,'' they meant ``physically motivated constraints'' more than statements that can be immune to question, rethink, and, if necessary, revise.\footnote{For a more extended discussion of von Neumann's axiomatics and their physical inputs, see \cite{redei_soft_2006}.}
+But, despite the undeniable importance of wanting to set physics on a firm mathematical footing, we must remember that by ``axioms,'' they meant ``physically motivated constraints'' more than statements that can be immune to question, rethink, and, if necessary, revise.\footnote{For a more extended discussion of von Neumann's axiomatics and their physical inputs, see \cite{redei_soft_2006, landsman_axiomatization_2022}.}
 
 Now, there are well-known shortcomings of Hilbert spaces, but they are deemed unproblematic: states are rays, not vectors; the Dirac delta is a distribution, not a function; and some elements from $L^2(\mathbb{R})$ don't seem apt candidates to be wave functions.
-Given the importance Hilbert spaces have had and the issues we have let slide, it's not a surprise that they have become fossilized as parts of axiomatic systems\footnote{Even axiomatic formulations of quantum field theory (see e.
-g.
-\cite{haag_local_1996}) that don't need to use Hilbert spaces to make meaningful claims about quantum-mechanical observables and states can reassure physicists unfamiliar with this approach by recovering the Hilbert spaces they know and love through the GNS construction.} without even considering whether we can do quantum mechanics with weaker assumptions.
+Given the importance Hilbert spaces have had and the issues we have let slide, it's not a surprise that they have become fossilized as parts of axiomatic systems\footnote{Even axiomatic formulations of quantum field theory \cite{haag_local_1996, horuzhy_introduction_1990} that don't need to use Hilbert spaces to make meaningful claims about quantum-mechanical observables and states can reassure physicists unfamiliar with this approach by recovering the Hilbert spaces they know and love through the GNS construction.} without even considering whether we can do quantum mechanics with weaker assumptions.
 So, despite all their success, perhaps it's time to ask if using Hilbert spaces to describe quantum states, an assumption whose physical motivation was meant to be revised as time passed, has become a dogma.
@@
 
-Let us first clarify the standard we believe should be usde to judge the physicality of a formal definition.
+Let us first clarify the standard we believe should be used to judge the physicality of a formal definition.
 Suppose we have a physical system under study in a given condition.
-We say a mathematical definition is \textbf{physical} if it is the proper mathematical characterization of the physical system.
+A mathematical definition is \textbf{physical} if it properly characterizes a physical system.
 That is, if it can be shown that the mathematical definition is needed to capture and only capture a particular aspect of the physical system.
@@
 
-For example, suppose we have a set of physically distinguishable objects, meaning that we can tell any two apart through an experimental test.
+For example, suppose we have a set of physically distinguishable objects, meaning that we can tell them apart through an experimental test.
 Experimentally, the best we can do is collect finite information in finite time.
-If we are given an arbitrarily long time, we can collect an arbitrary amount of finite information.
-An element of the set, then, must be identified by at most a countably infinite sequence of ones and zeros, so that, at some point, we would be able to tell the difference between the two sequences and therefore the objects.
-The set of countable sequences of zeros and one can be understood as the binary expansion of real numbers between zero and one, which has cardinality of the continuum.
-Therefore, any definition that would require a set of physical objects to have cardinality greater than the continuum, then, is \textbf{unphysical}.
-For example, non separable Hilbert spaces have cardinality greater than the continuum, and therefore should be considered unphysical.
+If given an arbitrarily long time, we can collect an arbitrary amount of finite information.
+An element of the set%TODO: Set of what?
+The example is clear, but this sentence is a bit unclear.
+, then, must be identified by, at most, a countably infinite sequence of ones and zeros so that, at some point, we can tell the difference between the two sequences and, therefore, the objects.
+The set of countable sequences of zeros and one can be understood as the binary expansion of real numbers between zero and one, which has the cardinality of the continuum.
+Therefore, any definition requiring a set of physical objects to have cardinality greater than the continuum is \textbf{unphysical}.
+For example, non-separable Hilbert spaces have cardinality greater than the continuum and, therefore, would be considered unphysical under this definition.
 
-Naturally, this does not mean that all sets with cardinality of the continuum or lower are physical.
-In general, testing physicality is harder as we need to check that all the mathematical features map to physical feature and no physical feature is left behind.
-Clearly, postulating that ``the state space of a quantum states has cardinality no grater than that of the continuum'' is certainly \textbf{physical} but, however, this tells us little.
-Let us then see what elements of Hilbert spaces should we consider physical in the context of quantum mechanics.
+Naturally, this does not mean all sets with a cardinality of the continuum or lower are physical.
+In general, testing physicality is harder as we need to check that all the mathematical features map to physical features and that no physical feature is left behind.
+Clearly, postulating that ``the state space of quantum states has cardinality no greater than that of the continuum'' is certainly \textbf{physical}, but this tells us little.
+Let us then see what elements of Hilbert spaces we should consider physical in the context of quantum mechanics.
+%TODO: I worry that this discussion on cardinality might be confusing since the "elements of Hilbert spaces" we'll discuss are parts of the definition of a Hilbert space, not the members of a set of physical objects.
 
@@
 Recall that a Hilbert space is a complete inner product space.
-That is, it is a vector product, with an inner product, closed under Cauchy sequences.
-We will look at the first two structures briefly, leaving all detail to the appendix, as their physical significance is not surprising.
+That is, it is a vector space with an inner product, and its Cauchy sequences converge under the norm defined by the inner product.
+We will look at the first two structures briefly, leaving all detail in the appendix, as their physical significance is unsurprising.
 
-Roughly speaking, a vector space is a set whose elements can be combined linearly: can be multiplied by constants and summed.
-If one regards superpositions as physical requirements of quantum theory, one may think that this structure is physical.
-However, the ability to interpret linear combination as superpositions requires the ability to keep track of orthonormal vectors.
-If we can only know whether two vectors are in the same ray, we are not able to recover the idea of a probability amplitude, as those are the coefficients of a linear combination between normalized and orthogonal states that return another normalized state.
-Therefore the vector space structure, by itself, is not physical.
+Roughly speaking, a vector space is a set whose elements can be combined linearly: they can be multiplied by constants and summed.
+If one regards superpositions as physical requirements of quantum theory, one may think this structure is physical.
+However, the ability to interpret linear combinations as superpositions requires keeping track of orthonormal vectors.
+If we can only know whether two vectors are in the same ray, we cannot recover the idea of a probability amplitude, as those are the coefficients of a linear combination between normalized and orthogonal states that return another normalized state.
+Therefore, the vector space structure, by itself, is not physical.
 
-If we add the inner product, we can define norm and orthogonality and therefore we can indeed express superpositions.
+If we add the inner product, we can define norm and orthogonality, and therefore we can indeed express superpositions.
 Additionally, the inner product structure also defines the Born rule, the transition probability during measurement.
-If one is comfortable in considering superpositions and the Born rule to be physical requirements of quantum mechanics, then the inner product structure is \textbf{physical}.
-However, some may not consider superpositions as actual physical objects, but more as mathematical constructs; others may object that the Born rule, in its standard interpretation, defines probability of transitions while a state space should be defined at equal time; others can argue that the states are defined by the rays, and therefore the requirements should only be in terms of those objects.
-For those and other reasons, we have an alternative, more necessary justification, which we include in the appendix, and can be summarized as follows.
+If one is comfortable considering superpositions and the Born rule to be physical requirements of quantum mechanics, then the inner product structure is \textbf{physical}.
+However, some may not consider superpositions as representing actual physical objects but more as mathematical constructs; others may object that the Born rule, in its standard interpretation, defines the probability of transitions while a state space should be defined at equal times; others can argue that the rays define the states, and therefore the requirements should only be in terms of those objects.
+For those and other reasons, we offer an alternative justification of the physicality of inner product vector spaces.
+Here is a summary, but the details are in the appendix.
 
 Given that, experimentally, no preparation or measurement is perfect, a physical theory must define the space of statistical ensembles.
-Mathematically, these form a convex set, that physically represents a set that allows statistical mixtures.
-The geometry of the space of statistical ensembles is uniquely defined by the inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtues.\footnote{There is a rich literature on Generalized Probability Theory in which people reconstruct Quantum Theory by imposing requirements on convex spaces.} For example, we can characterize superposition in the following way: a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ if and only if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
-That is, the ability to create superposition is conceptually equivalent to the ability to prepare the same mixture in different ways, another feature particular to quantum mechanics.
-Additionally, the inner product can be shown to be equivalent to determining the entropy of all equal mixtures of state pairs.
-These requirements are necessary in the sense that both classical and quantum mechanics need to keep track how ensemble combine under mixing and what entropy each ensemble have.
-
-The argument can be summed up as follows.
-On physical grounds, for any physical theory, we must be able to prepare statistical mixtures and characterize their entropy.
-The complex inner product space structure is exactly required to model the correct relationships of quantum mixtures, therefore the mathematical definition is necessitated by physical requirements.
-\textbf{The inner product space structure is physical}.
+Mathematically, these form a convex set that physically represents a set that allows statistical mixtures.
+The geometry of the space of statistical ensembles is uniquely defined by the inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.\footnote{There is a rich literature on Generalized Probability Theory in which people reconstruct Quantum Theory by imposing requirements on convex spaces.
+TODO: cite.} For example, we can characterize a superposition in the following way: a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ if and only if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
+That is, the ability to create a superposition is conceptually equivalent to the ability to prepare the same mixture in different ways, another feature particular to quantum mechanics.
+Additionally, the inner product can be shown to be equivalent to determining the entropy of all equal mixtures of pairs of states.
+Since we must be able to prepare statistical mixtures and characterize their entropy for any physical theory and the complex inner product space structure is exactly what is required to model the correct relationships of quantum mixtures, therefore \textbf{the inner product space structure is physical}.
 
@@
 Mathematically, this means that the space includes the limit of every Cauchy sequence; physically, this property is more difficult to characterize.
-We can look at an equivalent characterization:\cite{roman_2008}\footnote{Roman 2008, p.
-330 Theorem 13.8
-
-	Roman, Stephen (2008), Advanced Linear Algebra, Graduate Texts in Mathematics (Third ed.), Springer, ISBN 978-0-387-72828-5} the space is complete if, given a series of vectors
+We can look at an equivalent characterization \cite[Theorem 13.8, p.
+330]{roman_2008}.
+%\footnote{Roman 2008, p.
+330 Theorem 13.8	Roman, Stephen (2008), Advanced Linear Algebra, Graduate Texts in Mathematics (Third ed.), Springer, ISBN 978-0-387-72828-5}
+An inner product vector space is complete if, given a series of vectors
 $$ \sum _{k=0}^{\infty }u_{k}$$
 such that
-$$ \sum _{k=0}^{\infty }\|u_{k}\|<\infty \,$$
+$$ \sum _{k=0}^{\infty }\|u_{k}\|<\infty,$$
 the series converges in the space.
-This is now in terms of linear combinations, and therefore it is essentially telling us that we can prepare states that are superposition of infinitely many elements or, equivalently as we have seen, mixed states comprised of infinitely many pure states.
-Is this a physical requirement of just mathematical convenience?
+This characterization of completeness is in terms of linear combinations.
+It tells us that we can prepare states that are superpositions of infinitely many elements or, equivalently, as we have seen, mixed states comprised of infinitely many pure states.
+Is this a physical requirement or just mathematical convenience?
 
-Experimentally, we can imagining creating a superposition by splitting a beam along multiple paths, which is something typical, for example, in quantum optics.\footnote{TODO: cite} The above mathematical expression means splitting the original state into infinitely many progressively small components, and be able to place them where we please.
-However, it also assumes that we can place them over an infinite spatial range: we are literally saying that we can prepare states that are spread over infinite regions.
+Experimentally, we can create a superposition by splitting a beam along multiple paths, which is typical in quantum optics.\footnote{TODO: cite} The mathematical expression above means splitting the original state into infinitely many progressively small components and being able to place them where we please.
+However, that would assume that our experimental set-up spans an infinite region of space.
 This does not sound physical.
@@
 Again, given the difficulties of understanding superposition, let us reframe the result in terms of statistical quantities.
-Suppose that we are able to prepare pure states with arbitrary finite expectation value for position.
-Then completeness means we are also able to prepare pure states that falls off at infinity as $\frac{1}{x^2}$.
-The expectation value of $x$ would the integral of a function that falls off as $\frac{1}{x}$ which diverges.
+Suppose we can prepare pure states with arbitrarily finite expectation values for position.
+Then, completeness means we can also prepare pure states that fall off at infinity as $\frac{1}{x^2}$.
+The expectation value of $x$ would be proportional to the integral of a function that falls off as $\frac{1}{x}$, which diverges.
 Clearly, this does not make physical sense.
-When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that observed value can be arbitrarily large, not that we can literally spread values over an infinite range.\footnote{This is the difference between potential infinity and actual infinity.}
+When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that the observed value can be arbitrarily large, not that we can literally spread values over an infinite range.\footnote{This is the difference between potential infinity and actual infinity.
+Now, ``multiplying by $x$'' is not a well-defined operator in $L^2(\mathbb{R})$ unless we restrict the configuration space of our free particle to a compact interval of the real line.
+The fact that position operators must be bounded should recover some physicality, but that wouldn't be a property of the states themselves, which is what we're interested in studying.}
 
@@
 On physical grounds, we assume the existence of quantities, like position, that can take arbitrarily large values.
-This requires that for each possible value there must exist some state whose expectation matches that value.
-Requiring completeness mathematically forces us to posit states with an infinite expectation value, which cannot be implemented physically.
-The mathematical definition implicitly assumes objects that cannot be physically realized.
-\textbf{The use of completeness for the space of quantum systems in unphysical.}
+This requires that, for each possible value, there must exist some state whose expectation matches that value.
+Requiring completeness mathematically forces us to posit states with an infinite expectation value, but that cannot be implemented physically.\footnote{Philosophers interested in perturbative quantum field theory have formulated interpretive strategies to make sense of some of the divergent series that arise, but surveying them would take us too far afield.
+For an example, see \cite{miller_mathematical_2023}.} The mathematical definition implicitly assumes objects that cannot be physically realized, so \textbf{the use of completeness for the space of quantum systems in unphysical.}
 
@@
 
-It should not be surprising that closure over infinite sequences is problematic, as properly handling infinity is one of the key problems in mathematics.\footnote{Weyl stated ``Mathematics is the science of the infinite, its goal the symbolic comprehension of the infinite with human, that is finite, means.'' Hermann Weyl, "The Open World: Three Lectures on the Metaphysical Implications of Science," (1932) as quoted in Mind and Nature: Selected Writings on Philosophy, Mathematics, and Physics (2009) ed.
+It should not be surprising that closure over infinite sequences is problematic, as properly handling infinity is one of the key problems in mathematics.\footnote{Weyl stated, ``Mathematics is the science of the infinite, its goal the symbolic comprehension of the infinite with human, that is finite, means.'' Hermann Weyl, "The Open World: Three Lectures on the Metaphysical Implications of Science," (1932) as quoted in Mind and Nature: Selected Writings on Philosophy, Mathematics, and Physics (2009) ed.
 Peter Pesic.
-Several branches of mathematics were developed to handle infinities, both large and small.
+Several branches of mathematics were developed to handle large and small infinities.
 } We should understand the impact on the physics of using completeness for that goal.
+%TODO: This sentence was a bit difficult to parse.
 
-If an inner product space is finite dimensional, it is automatically complete.\footnote{This is exactly why, mathematically, completeness is nice as it makes infinite dimensional space behave like finite dimensional ones.} If the space is not complete, we can always complete it by enlarging the space.
-The prevalent attitude among physicists is that this pose no problem: we simply have a bigger set of objects than what we strictly need, and we simply discard what we do not need.
-Unfortunately, having a bigger space often creates bigger problems.
+If an inner product space is finite-dimensional, it is automatically complete.\footnote{This is exactly why completeness is mathematically desirable: infinite-dimensional spaces behave like finite-dimensional ones.} If the space is not complete, we can always complete it by enlarging it.
+The prevalent attitude among physicists is that this does not pose a problem: we simply have a bigger set of objects than we strictly need, and we simply discard what we do not need.
+However, having a bigger space with redundant elements often creates bigger problems.
 
-For example, take the problem of defining volumes in three dimensional euclidean space $\mathbb{R}^3$.
+For example, take the problem of defining volumes in three-dimensional Euclidean space $\mathbb{R}^3$.
 Formally, we want to find a measure $\mu(U)$ that returns the size of the region $U$.
@@
 Clearly, this does not make physical sense.
-However, if we restrict ourselves with a measure on the Borel sets instead of all sets, we get a consistent definition.
-Therefore, in mathematics, bigger is not always better:\footnote{In another work, we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
-In many cases, we have found that physical requirements naturally map to well behaved spaces.} we need to look more closely at what the math dragged in.
+However, we get a consistent definition if we restrict ourselves to a measure on the Borel sets instead of all sets.
+Therefore, bigger is not always better in mathematics:\footnote{In another work TODO: cite, we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
+We have often found that physical requirements naturally map to well-behaved spaces.%TODO: Is this related to the spectral theorem somehow?
+} we need to look more closely at what the math dragged in.
 
-Unitary transformation play an important role in the theory of Hilbert spaces for at least two reasons.
+Unitary transformations play an important role in the theory of Hilbert spaces for at least two reasons.
 The first is that two Hilbert spaces that differ by a unitary transformation are mathematically equivalent.
-A change of representation, like a change of basis, is a unitary transformation.
+Like a change of basis, a change of representation is a unitary transformation.
 The second reason is that time evolution for an isolated system is represented by a unitary transformation.
-Let us see, then, the interplay between unitary transformations and our previous findings.
+Let us see, then, the interplay between unitary transformations and our previous findings.\footnote{More precisely, the first reason only holds if the assumptions of the Stone-von Neumann theorem hold.
+For the conditions and physical systems under which those assumptions are unwarranted and the interpretive problems of trying to understand physical equivalence for systems that are not unitarily equivalent, see \cite{ruetsche_interpreting_2011}.
+The second reason is a consequence of the Stone theorem mentioned in the introduction.
+Notice that the definition of unitary transformations requires the inner product.
+Yet another reason for its physicality!}
 
@@
 Therefore, there will be unitary transformations that map states with finite expectation values to infinite ones.
-This means that whether something has a finite expectation value or not will depend on the choice of representation.
+This means that whether something has a finite expectation value will depend on the choice of representation.
 As a concrete example, consider the following two wave functions
@@
 \end{align}
-Both states are symmetric, and therefore will have zero expectation value for $x$.
-The first distribution is a Gaussian, and therefore all moments of the distribution, the expectation for $x^n$, are finite.
+Both states are symmetric and, therefore, will have zero expectation value for $x$.
+The first distribution is a Gaussian; therefore, all moments of the distribution, the expectation for $x^n$, are finite.
 The second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all moments above the first.
@@
 \end{equation}
-Therefore, through a variable change, we can change a state that has finite expectation value for all moments greater than the first into a state that has infinite expectation value for the same moments.
-Variable changes, in terms of the Hilbert space, are unitary transformation.
-Therefore the theory of Hilbert spaces, as applied to quantum mechanics, would tell us that the two representations are equivalent.
+Therefore, through a variable change, we can change a state with a finite expectation value for all moments greater than the first into a state with an infinite expectation value for the same moments.
+Variable changes are unitary transformations on the Hilbert space.
+%TODO: I worry that I'm not sure if this change of variables is not implementable by unitary operators...writing down the explicit unitary transformation would be great.
+Therefore, the theory of Hilbert spaces, as applied to quantum mechanics, would tell us that the two representations are equivalent.
 
 On physical grounds, these are clearly not equivalent.
-Identifying a state through expectation value is a common technique (i.e quantum tomography).
-If a unitary transformation can change finite expectations to infinite ones, we are effectively changing what can be distinguished experimentally.
-Not only we cannot measure infinite values, but different finite values would get lumped into the same infinite value.
-In the above example, if we increased the variance, one observer will see all the even moments change, yet the other would see them all equally infinite.
-The math, as it stands, lumps together physically inequivalent representations into the same equivalence class, making it impossible later to separate physical problem from unphysical ones.
+Reconstructing a state from its expectation values is a common technique, quantum tomography.
+If a unitary transformation can change finite expectations to infinite ones, we effectively change what can be distinguished experimentally.
+Not only can we not measure infinite values, but different finite values would get lumped into the same infinite value.
+%TODO: The comment on tomography made me think that tomographic procedures also have an uncertainty associated with them (some collaborators and I have a paper on this), so this would mean that we have a device with infinitely good resolution, which is nonsense!
+If this argument seems right, we can include it.
+In the example above, if we increased the variance, one observer would see all the even moments change, but the other would see them all equally infinite.
+The math, as it stands, takes physically inequivalent states as equivalent, making it impossible later to separate physical problems from unphysical ones.
 
@@
 \end{equation}
-This is clearly continuous in $t$ and, for $t=0$, we have the identity transformation and therefore, mathematically, there is no reason why this wouldn't be a valid time evolution.
-What this does is stretching back and forth the wavefunction, such that the state keeps oscillating between the above two functions, making the expectation value oscillate from finite to infinite in finite time.
+This is clearly continuous in $t$ and, for $t=0$, we have the identity transformation, and therefore, mathematically, there is no reason why this wouldn't be a valid time evolution.
+%TODO: This would be more convincing by showing the unitary transformation reflecting the change of variables.
+What this does is stretch back and forth the wavefunction, such that the state keeps oscillating between the two functions above, making the expectation value oscillate from finite to infinite in finite time.
 This, again, is physically untenable.
 
-Again, most of the time these issues do not arise when performing actual calculations because the physics makes us work in a corner of the structure where things are reasonable.
-The issue is that when one tries to prove theorems or general results on quantum theory, one currently is forced to consider the entire Hilbert space.
-It is unreasonable to expect that all theorems valid on all physical cases would have valid generalizations on unphysical ones, and therefore we are likely missing useful results.
-We therefore wonder whether many of the technical issues we have in quantum theory (e.g.
-non existence of a valid measure for path integrals, inequivalent representation of interacting theories, non existence of the categorial tensor product for Hilbert spaces, having to deal with infinities and infinite correlations)\footnote{TODO: cite} are simply a consequence of the use, at a fundamental level, of unphysical objects.
+Again, these issues often do not arise when performing actual calculations because physics makes us work in the corner of the structure where things are reasonable.
+The issue is that when one tries to prove theorems or general results on quantum theory, one is currently forced to consider the entire Hilbert space.
+It is unreasonable to expect all theorems valid in all physical cases to have valid generalizations for unphysical ones.
+We are likely missing out on useful results by not considering what is provable within the subset of physical cases.
+We, therefore, wonder whether many of the technical issues we have in quantum theory (e.
+g., the non-existence of a valid measure for path integrals, inequivalent representations for interacting theories, the non-existence of the categorical tensor product for Hilbert spaces, etc.)\footnote{See \cite[\S A.4]{glimm_quantum_1987} for the non-existence of infinite products of Lebesgue measures.
+Glimm and Jaffe also mention a theorem by Minlos (3.4.2) suggesting that path integral measures \textit{can} be defined in Schwartz spaces that, as far as we know, has not been paid attention to in the literature.
+The locus classicus for Haag's theorem and the interaction picture in quantum field theory is still \cite{earman_haags_2006}.
+For categorical tensor products, see TODO.} having to deal with infinities and infinite correlations are simply a consequence of using unphysical objects at a fundamental level.
 
@@
 
-While we do need infinite dimensional spaces to handle unbound quantities, such as number of particles, or continuous quantities, such as position, it seems that Cauchy completeness is not the proper characterization.
-While finding the proper one goes beyond the scope of this work, we do want to show that alternatives are possible.
+While we do need infinite-dimensional spaces to handle unbound quantities, such as the number of particles, or continuous quantities, such as position, it seems that Cauchy-completeness is not the proper characterization.
+While finding the proper one goes beyond the scope of this work, we want to show that alternatives are possible.
 
-A reasonable position would be to require the expectation of all polynomial of position and momentum to be finite.
+A reasonable alternative would require the expectation of all polynomials of position and momentum to be finite.
 In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m - P^mX^n)$ to be finite.
-This restricts our space to those functions that are infinitely smooth (i.e.
-all derivative exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.
+This restricts our space to those functions that are infinitely smooth (i.
+e., all derivatives exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.
 
-The Schwartz space has a number of other interesting properties, explored in a bit more detail in the appendix. For example:
+The Schwartz space has several other interesting properties, explored in more detail in the appendix. For example:
+
 \begin{itemize}
-	\item it is fully identified by the expectations - two Schwartz are the same if and only if all expectations are the same; quantum tomography on all polynomials fully reconstructs the state
+\item the expectations fully identify it - two Schwartz are the same if and only if all expectations are the same; quantum tomography on all polynomials fully reconstructs the state
 \item it is inner product space with the same norm of $L^2$
-	\item it is a dense subset of $L^2$ Hilbert space - any element of the Hilbert space can be approximated by a Schwartz function with arbitrary level of precision
-	\item it is complete with respect to the expectations - a sequence of functions for which all expectations of all polynomials of position and momentum converge will converge in the Schwartz space  
+\item it is a dense subset of $L^2$ Hilbert space - any element of the Hilbert space can be approximated by a Schwartz function with an arbitrary level of precision\footnote{Even the problematic position operators can be defined starting from Schwartz spaces \cite[Threorem 5.23]{moretti_spectral_2017}.}
+\item it is complete with respect to the expectation values - a sequence of functions for which all expectation values of all polynomials of position and momentum converge will converge in the Schwartz space
 \item its closure over all Cauchy sequences recovers the Hilbert space - the only thing missing are the Cauchy sequences for which the limit of at least one expectation is not well defined
@@
 \end{itemize}
-These properties make the Schwartz space a much reasonable candidate to capture the physics.
-Convergence over vectors in an abstract space is replaced by converges of expectation values of actual observables.
-The point is that alternatives exist and should be fully explored.
+These properties make the Schwartz space a much more reasonable candidate to capture the physics.
+The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} The point is that alternatives exist and should be fully explored.
 
@@
 
-We have made the case that Hilbert spaces in quantum mechanics are \textbf{unphysical} as they necessarily bring in states with infinite expectation values, coordinate transformations that turn finite expectations in infinite ones, and unitary transformations that do the same in finite time.
-Theorems that are valid on Hilbert spaces must be valid over these unphysical cases, meaning that we are possibly missing results and techniques that would be physically interesting.
+We have made the case that Hilbert spaces in quantum mechanics are \textbf{unphysical} as they necessarily bring in states with infinite expectation values, coordinate transformations that turn finite expectations into infinite ones, and unitary transformations that do the same in finite time.
+Theorems that are valid on Hilbert spaces range over these unphysical cases, meaning that we are possibly missing results and techniques that would be physically interesting.
 
-The more general problem is that the physics community has become complacent in simply accepting mathematical definitions, without properly understanding their limit of applicability for physical theories.
+The more general problem is that the physics community has become complacent in simply accepting mathematical definitions without properly understanding their limit of applicability to physical theories.
 We believe that meaningful progress on the foundations of physics cannot happen without understanding the implicit assumptions embedded in the most basic mathematical tools.
@@
 
-In this appendix we provide the same arguments presented in the main body of the article expanded with the full mathematical detail.
+This appendix provides the same arguments in the article's main body, but expanded with full mathematical detail.
 
@@
 The first task is to show that the vector space structure is not, by itself, physical.
-Since quantum states are represented by rays of a Hilbert space, and since rays can be defined simply on top of the vector space structure alone, one may think that the vector space structure, by itself, is physical.
-The issue is that, without an inner product, one cannot make physical sense of the coefficients one uses for linear combinations, therefore superpositions becomes ill-defined.
+Since rays of a Hilbert space represent quantum states, and rays can be defined simply on top of the vector space structure alone, one may think that the vector space structure, by itself, is physical.
+The issue is that, without an inner product, one cannot make physical sense of the coefficients one uses for linear combinations, and, therefore, superpositions become ill-defined.
 
-To show the problem, we prove the following proposition that tells us that there are infinitely many ways to decompose a vector representing a state in terms of vectors that represent two other states, meaning that the coefficients of the linear combination are arbitrary.
+To show the problem, we prove the following proposition that tells us that there are infinitely many ways to decompose a vector representing a state in terms of vectors representing two other states, meaning that the coefficients of the linear combination are arbitrary.
 \begin{prop}\label{vector_insufficient}[The vector space structure is not enough to uniquely identify superpositions.] Let $\mathcal{H}$ be a complex vector space and let $\textbf{P}(\mathcal{H})$ be its projective space.
@@
 \end{equation}
-This concludes the proof.
 \end{proof}
@@
 The intuition that a state has a single linear decomposition in terms of other vectors requires, first of all, that the vectors are normalized.
-To identify the coefficient as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires the decomposition to be done on orthogonal vectors.\footnote{Mathematically, this means turning the vectors $\{\phi_i\}$ into an orthonormal set using Gram-Schmidt orthonormalization, which requires the inner product.} That is, linear combinations require the inner product to be understood physically as superpositions.
-The vector space structure by itself is therefore insufficient to provide a clear connection to physics.
+To identify the coefficient as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires decomposition on orthogonal vectors.\footnote{Mathematically, this means turning the vectors $\{\phi_i\}$ into an orthonormal set using Gram-Schmidt orthonormalization, which requires the inner product.} That is, linear combinations require the inner product to be understood physically as superpositions.
+Therefore, the vector space structure is insufficient to provide a clear connection to physics.
 
-We could now use the above reasoning to argue that the inner product space structure, as a whole, is physical, because it indeed characterizes both superpositions and the Born rule.
-These are features required by quantum mechanics, and therefore necessary and physically motivated.
-Many will find this argument convincing enough, however there a couple of objective problems.
+We could now use the above reasoning to argue that the inner product space structure, as a whole, is physical because it indeed characterizes both superpositions and the Born rule.
+These are features required by quantum mechanics and, therefore, representationally necessary and physically motivated.
+Many will find this argument convincing enough, but there are several problems.
 
 First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs.
-This is an area where different interpretations may differ,\footnote{TODO: add citations} and the physicality of a mathematical representation should not be interpretation dependent.
-Second, the Born rule, in its standard characterization, brings in one of the most controversial aspects of the theory, measurements, which suffers from an interpretation problem as well.
-Lastly, the inner product is critical for both the Born rule, which is a transition probability of transition before and after the measurement, and linear decomposition into superposition, which is done at equal time.
-The goal, then, is to find a characterization of our physical requirement that is interpretation independent, more cohesive and perhaps more physically necessary.
+This is an area where different interpretations may differ \cite{albert_quantum_1994, wallace_everett_2013, howard_complementarity_2021}, and the physicality of a mathematical representation should not be interpretation-dependent.
+Second, in its standard characterization, the Born rule brings in one of the most controversial aspects of the theory, measurements, which suffers from an interconnected interpretation problem.
+Lastly, the inner product is critical for both the Born rule, which is a transition probability of transition before and after the measurement, and linear decomposition into superposition, which is done at equal times.
+The goal, then, is to find a characterization of our physical requirement that is interpretation-independent, more cohesive, and perhaps more physically necessary %TODO: Not sure what ``necessary'' means here.
 
@@
 
-The idea is to link superposition to another feature of quantum mechanics: the non-unique decomposition of mixed state into pure states.
-For example, if we take the maximally mixed state for a spin 1/2 system, this can be implemented with an equal mixture of spin up and spin down, or with an equal mixture of spin left and spin right.
+The idea is to link superposition to another feature of quantum mechanics: the non-unique decomposition of mixed states into pure states.
+For example, if we take the maximally mixed state for a spin 1/2 system, this can be implemented with an equal mixture of spin up and spin down or with an equal mixture of spin left and spin right.
 The idea is that a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ exactly because there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
@@
 Let $\{\psi_i\}, \{\phi_j\} \in \mathcal{H}$ be two finite sets of normalized vectors.
-Suppose there exists a density operator that can be expressed as a strict convex combination of either set, that is we can find two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}_{>0}$ such that
+Suppose there exists a density operator that can be expressed as a strict convex combination of either set.
+That is, we can find two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}_{>0}$ such that
 \begin{enumerate}
@@
 If not, the two elements will identify a line, which must intersect the boundary of the convex set in two places.
-One must be $\psi$, since $\psi$ is a pure state and therefore an extreme point.
+One must be $\psi$ since $\psi$ is a pure state and, therefore, an extreme point.
 Let $\rho_1$ be the other point.
@@
 That region will be a convex set constrained on a subspace $V_1$ of $V$.
-Note that $V_1$ does not include $\psi$ and it does not span the direction identified by $\rho_1$ and $\psi$.
+Note that $V_1$ does not include $\psi$, and it does not span the direction identified by $\rho_1$ and $\psi$.
 This means that we can't have $\rho_1 = \rho$ as $\<\psi|\rho|\psi\>\neq 0$.
@@
 Let $\mathcal{H}$ be a complex vector space.
-Let $\psi \in \mathcal{H}$ be a normalized vector and $\{\phi_j\} \in \mathcal{H}$ a finte sets of normalized vectors that does not contain $\psi$.
+Let $\psi \in \mathcal{H}$ be a normalized vector and $\{\phi_j\} \in \mathcal{H}$ a finite set of normalized vectors that does not contain $\psi$.
 Then $\psi$ is a superposition of $\{\phi_j\}$ if and only if there exists a density operator $\rho$ that can be expressed both as a strict convex combination of $\{\phi_i\}$ and of another set of normalized vectors that contain $\psi$.
@@
 \begin{proof}
-	Suppose $\psi$ is a normalized state that can be written as $\psi = \sum_{i=1}^n c_i \phi_i$ where $\{\phi_i\} \in \mathcal{H}$ is a finite sets of normalized vectors that do not contain $\psi$.
+Suppose $\psi$ is a normalized state that can be written as $\psi = \sum_{i=1}^n c_i \phi_i$ where $\{\phi_i\} \in \mathcal{H}$ is a finite set of normalized vectors that does not contain $\psi$.
 Since $\{\phi_j\}$ are not necessarily orthogonal, it is not necessarily true that $|\<\psi|\phi_i\>| = |c_i|^2$ or that $|\<\psi|\phi_i\>| \neq 0$ for all $i$.
@@
 
-	Conversely, suppose we have a density operator $\rho$ that can be expressed both as a strict convex combination of $\{\phi_i\}$ and of another set of normalized vectors that contain $\psi$.
+Conversely, suppose we have a density operator $\rho$ that can be expressed as a strict convex combination of $\{\phi_i\}$ and another set of normalized vectors containing $\psi$.
 By Proposition \ref{prop_densitySpan} the span of $\{\phi_i\}$ will include $\psi$ and therefore $\psi$ is a superposition of $\{\phi_i\}$.
@@
 As for the inner product, we show that the Born rule can be understood as characterizing the entropy of equal mixtures over state pairs.
-This gives us a geometric understanding of the inner product at equal time, instead of a probability of transition, which brings it more in line with the use during superposition.
+This gives us a geometric understanding of the inner product at equal times instead of a probability of transition, which brings it more in line with the use during superposition.
 It should also be noted that the von Neumann entropy of a mixture can be understood as the minimum Shannon entropy over all possible decompositions,\footnote{TODO: cite} which again shows the connection between the geometry of the convex space of statistical ensembles and the inner product.
@@
 \end{equation}
-Then $p$ and $I$ can be reconstructed from other. That is, there exists an invertible function $f$ such that
+Then $p$ and $I$ can be reconstructed from each other. That is, there exists an invertible function $f$ such that
 \begin{equation}
@@
 \begin{proof}
-	Copy proof from other paper.
+TODO: Copy proof from other paper.
 \end{proof}
 
-Having shown the connection between superposition and non-unique decomposition of mixed state, and between the Born rule and the entropy of mixtures, we can now use these to justify the physicality of the complex inner product space structure for quantum states.
+Having shown the connection between superposition and non-unique decompositions of mixed states and between the Born rule and the entropy of mixtures, we can now use these to justify the physicality of the complex inner product space structure for quantum states.
 \begin{prop}
@@
 \begin{justification}
-	Since physical preparations and measurements are most of the time not perfect, a physical theory must define statistical ensembles.
-Part of that definition includes keeping track of statistical ensembles under statistical mixtures, that is which particular mixture is obtained by mixing particular states in particular proportions.
-A particular feature of quantum mixed states is that, unlike classical states, they do not in general provide a unique decomposition in terms of pure states.
+Since physical preparations and measurements are often not perfect, a physical theory must define statistical ensembles.
+Part of that definition includes keeping track of statistical ensembles under statistical mixtures, that is, which particular mixture is obtained by mixing particular states in particular proportions.
+A particular feature of quantum mixed states is that, unlike classical states, they do not generally provide a unique decomposition in terms of pure states.
 This non-unique decomposition, as shown by proposition \ref{prop_superpositionIsDecomposition}, is equivalent to the ability to create superpositions of pure state.
-The inner product structure is therefore the exact structure needed to characterize quantum ensembles and their properties, which is required by the physics.
+Therefore, the inner product structure is the exact structure needed to characterize quantum ensembles and their properties, which is required by the physics.
 
-	Additionally, a physical theory must be able to define the entropy of each ensemble, as required by thermodynamics.
-In quantum mechanics, the thermodynamic entropy is recovered by the von Neumann entropy, which Proposition \ref{prop_innerProductIsEntropy} show it is exactly captured by the inner product.
+Additionally, a physical theory must be able to define each ensemble's entropy, as thermodynamics requires.
+In quantum mechanics, the thermodynamic entropy is recovered by the von Neumann entropy, which Proposition \ref{prop_innerProductIsEntropy} shows is exactly captured by the inner product.
 This again shows that the inner product structure is required to capture a necessary physical requirement.
@@
 
-The Hellinger–Toeplitz theorem states that a self-adjoint operator $O$ defined on the whole Hilbert space $\mathcal{H}$ is necessarily bound.
+The Hellinger–Toeplitz theorem states that a symmetric operator $O$ defined on the whole Hilbert space $\mathcal{H}$ is necessarily bound.
 Therefore, an unbound operator $X$ cannot be defined on the whole space, meaning that the operator must diverge or be ill-defined on a subset of $\mathcal{H}$.
@@
 \begin{prop}
-	Let $\mathcal{H}$ be a Hilbert space and $X : D(X) \to Y$ a unbounded operator where $D(X)$ is the domain of $X$ and $Y$ is another Hilbert space.
+Let $\mathcal{H}$ be a Hilbert space and $X : D(X) \to Y$ an unbounded operator where $D(X)$ is the domain of $X$, and $Y$ is another Hilbert space.
 Then $D(X)$ is an inner product space that is not complete.
@@
 Therefore $D(X)$ is an inner product space.
-It cannot be complete by the converse of the Hellinger–Toeplitz theorem.
+It cannot be completed by the converse of the Hellinger–Toeplitz theorem.
 \end{proof}
@@
 The operator is unbounded because the position can be arbitrarily large.
-We can therefore construct a Cauchy sequence $\{\psi_i\}$ such that $\lim\limits_{i \to \infty}||X\psi|| = +\infty$, that is take a sequence of vectors for which the expectation of position is larger and larger.
+We can therefore construct a Cauchy sequence $\{\psi_i\}$ such that $\lim\limits_{i \to \infty}||X\psi|| = +\infty$, that takes a sequence of vectors for which the expectation of position is larger and larger.
 Because of completeness, the limit of the Cauchy sequence will be in the Hilbert space.
@@
 
-The matter would still be manageable if we had a way to keep these degenerate cases isolated, but this is not the case. In general, a unitary operator can move vectors in and out of the domain.
+The matter would still be manageable if we had a way to keep these degenerate cases isolated, but this is not the case. Generally, a unitary operator can move vectors in and out of the domain.
 \begin{prop}
-	Let $\mathcal{H}$ be a Hilbert space, $X : D(X) \to Y$ a unbounded operator and $U : \mathcal{H} \to \mathcal{H}$ a unitary operator. Then, in general, $U(D(X)) \notin D(X)$.
+Let $\mathcal{H}$ be a Hilbert space, $X : D(X) \to Y$ an unbounded operator, and $U : \mathcal{H} \to \mathcal{H}$ a unitary operator. Then, in general, $U(D(X)) \notin D(X)$.
 \end{prop}
@@
 Then we can construct a unitary operator $U$ such that $\phi = U(\psi)$.
-For example, we can take the operator that performs a rotation in the subspace identified by $\psi$ and $\phi$ and leaves the orthogonal subspace unchanged.
+For example, we can take the operator that performs a rotation in the subspace identified by $\psi$ and $\phi$, leaving the orthogonal subspace unchanged.
 We can now take $\psi$ to be in $D(X)$ while $\phi$ to not be in $D(X)$.
-Therefore, in general, a unitary operator can move vectors in and out the domain of an unbounded operator.
+Therefore, in general, a unitary operator can move vectors in and out of the domain of an unbounded operator.
 \end{proof}
@@
 
-More concretely, a change of variable $X$ to $Y$ would be a unitary operator as is just a change of representation.
+More concretely, a change of variable $X$ to $Y$ would be a unitary operator as it is just a change of representation.
 However, we can construct changes of variables such that the domain of $X$ is not exactly mapped to the domain of $Y$.
@@
 \end{equation}
-To find a coordinate transformation that take finite expectations to infinite expectation, we set the initial and final wavefunctions as follow:
+To find a coordinate transformation that takes finite expectations to infinite expectation, we set the initial and final wavefunctions as follow:
 \begin{align}
@@
 \end{align}
-The first distribution is a Gaussian, and therefore all moments of the distribution, the expectation for $x^n$, are finite.
+The first distribution is a Gaussian; therefore, all moments of the distribution, the expectation for $x^n$, are finite.
 The second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all moments above the first.
@@
 The graph of the above function increase so fast that, when plotted, it looks like it has a vertical asymptote.
-Nonetheless, it is an invertible function over the whole domain, and will induce a unitary transformation on the Hilbert space.\footnote{Note that we could implement a similar coordinate transformations in classical mechanics, which would also change finite moments of a distribution into infinite ones.
+Nonetheless, it is an invertible function over the whole domain and will induce a unitary transformation on the Hilbert space.\footnote{Note that we could implement a similar coordinate transformation in classical mechanics, which would also change finite moments of a distribution into infinite ones.
 It may be that, like differentiable manifolds are constrained to only differentiable coordinate transformation, we need ``statistical manifolds'' that are constrained to coordinate transformations that preserve the finiteness of moments of distributions.
@@
 \end{equation}
-The idea is that the evolution will simply stretch and shrink the wavefunction intermittently, potentially oscillating expectation values from finite to infinite and viceversa in finite time.
+The idea is that the evolution will simply stretch and shrink the wavefunction intermittently, potentially oscillating expectation values from finite to infinite and vice versa in finite time.
 Clearly, this is not a meaningful time evolution, yet it will correspond to a strongly continuous one-parameter unitary group and, by Stone's theorem, will even admit a Hamiltonian.
 
-The mathematician that points out that this all known, and there is nothing new, is missing the point.
-The physicists that notes that, yes, those elements are clearly unphysical, but I don't use them, so I do not care, is also missing the point.
-The point is that, physically, we will want to make general statements such as ``for all observer X is valid'' or ``under all evolution Y is valid''.
-Given that Hilbert spaces mix physically clearly meaningful object with clearly unphysical ones, we have no way to make those statements precisely.
+The mathematician that points out that this is all known and there is nothing new is missing the point.
+The physicists that note that ``yes, those elements are clearly unphysical, but I don't use them, so I do not care,'' are also missing the point.
+The point is that, physically, we will want to make general statements such as ``for all observers, X is valid'' or ``under all evolution, Y is valid.'' Given that Hilbert spaces mix physically clearly meaningful objects with clearly unphysical ones, we have no way to make those statements precisely.
 
@@
 \begin{prop}
-	Infinite dimensional complex Hilbert spaces are unphysical when used to represent the state space of quantum systems due to completeness.
+Infinite-dimensional complex Hilbert spaces are unphysical when used to represent the state space of quantum systems due to completeness.
 \end{prop}
 \begin{justification}
-	If a quantum system is modeled by a finite dimensional space, and since every finite dimensional space is complete, finite dimensional complex Hilbert spaces are physical.
+If a finite-dimensional space models a quantum system, and since every finite-dimensional space is complete, finite-dimensional complex Hilbert spaces are physical.
 
-	Some physical system that are described by unbounded discrete quantities (e.g.
-number of particles) or by continuous quantities (e.g.
-position) which require infinite dimensional spaces.
-Infinite dimensional complex inner product spaces, then, are needed to represent quantum system.
+Some physical systems are described by unbounded discrete quantities (e.
+g., number of particles) or continuous quantities (e.
+g., position), which require infinite-dimensional spaces.
+Infinite-dimensional complex inner product spaces, then, are needed to represent quantum systems.
 
 An unbounded operator cannot be defined on all the elements of a Hilbert space by the converse of the Hellinger-Toeplitz theorem.
-This means that if represents states as rays on a Hilbert space and observables as operators, the unbounded observable will not be well defined, not even as a statistical quantity, on all states.
-In particular, whether the statistical properties of an observable are well defined, whether it is possible to experimentally identify a state through tomography, would be observer dependent.
+This means that if one represents states as rays on a Hilbert space and observables as operators, the unbounded observable will not be well defined, not even as a statistical quantity, on all states.
+In particular, whether the statistical properties of an observable are well defined, whether it is possible to identify a state through tomography experimentally would be observer-dependent.
 Moreover, whether the expectation of an observer is finite or not would be something that can change in finite time.
 
-	On physical ground, we must have that an observable has well defined statistical properties over all states, that this property is covariant and that is preserved by time evolution.
-Hilbert spaces do not satisfy this requirement and are therefore unphysical.
+On physical ground, we must have that an observable has well-defined statistical properties over all states, that this property is covariant, and that it is preserved by time evolution.
+Hilbert spaces do not satisfy this requirement and are, therefore, unphysical.
 \end{justification}
@@
 \item complex inner product spaces are needed
-	\item infinite dimensional spaces are needed
+\item infinite-dimensional spaces are needed
 \item completeness is unphysical.
@@
 In other words, we need a stricter way to characterize convergence within the state space so that we can manage the infinite dimensionality.
-Conceptually, we need to treat the infinities coming from unbound quantities as potential infinities, not actual infinities: we never actually have infinitely many particles or a system positioned at infinity, we have an infinite range of possible values.
+Conceptually, we need to treat the infinities coming from unbound quantities as potential infinities, not actual infinities: we never actually have infinitely many particles or a system positioned at infinity; we have an infinite range of possible values.
 
@@
 Suppose we have a distribution $\rho(x)$ over position $x$.
-We would be requiring the expectation of any function to be finite, which would include $e^x$, $e^{e^x}$, $e^{e^{e^x}}$ and so on.
+We would require the expectation of any function to be finite, which would include $e^x$, $e^{e^x}$, $e^{e^{e^x}}$, and so on.
 The only way this can work is if the function has compact support.
 Physically, this seems too restrictive as it excludes functions like normal distributions.
-At a closer investigation, the Paley-Wiener theorem\footnote{TODO: find citation} tells us that the Fourier transform of a function with compact support is an entire function, and entire function do not have compact support.\footnote{TODO: find citation} Therefore, we cannot have compact support in both position and momentum.
+At a closer investigation, the Paley-Wiener theorem\footnote{TODO: find citation--Maybe Prop.
+3.115 in \cite{moretti_spectral_2017}?} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\footnote{TODO: find citation} Therefore, we cannot have compact support in both position and momentum.
 \footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
@@
 The only way out, then, is to require only a certain class of observables to have finite expectation.
-Physically, it would make sense to require exactly those observables that are needed to fully reconstruct the state through quantum tomography.
-A good set in that respect seems to be the set of polynomial of conjugate quantities, like position and momentum.
+Physically, it would make sense to require exactly those observables needed to reconstruct the state through quantum tomography fully.
+A good set in that respect seems to be the set of polynomials of conjugate quantities, like position and momentum.
 In quantum mechanics, this means requiring all expectations of $X^nP^m = -\imath \hbar x^n\partial_x ^m$ to be finite.
-This restricts our space to those functions that are infinitely smooth (i.e.
-all derivative exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions.
-Requiring finite expectation for all polynomials of position and momentum, then, does not give us the usual $L^2$ Hilbert space, it gives us the Schwartz space.
+This restricts our space to those functions that are infinitely smooth (i.
+e., all derivatives exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions.
+Requiring finite expectation for all polynomials of position and momentum, then, does not give us the usual $L^2$ Hilbert space; it gives us the Schwartz space.
 
-Not only the expectations of all polynomials exist, but they are enough to identify the elements and to define limits.
-The Schwartz space is in fact a Fr\'echet, meaning that is a Hausdorff space, with the topology induced by the seminorms given by the expectations and it is complete with respect to those seminorms.
-Physically, this means that measurements on the polynomials (i.e.
-the topology of the seminorms) are enough to identify the states (i.e.
-the singletons, which are closed sets as per Hausdorff).\footnote{This is not true in classical mechanics.
-Even in the single variable case, there are distinct distributions that have the same moments.
-However, if we assume compact support, distinct distributions have distinct moments even in the multi variable case.\cite{moment_problem_2017} It is curious that the moment problem in quantum mechanics works differently.} Moreover, if we have a sequence for which all expectations converge, we are guaranteed to converge to another state (i.e.
-completeness with respect to the seminorms).
+Not only do the expectations of all polynomials exist, but they are enough to identify the elements and define limits.
+The Schwartz space is, in fact, a Fréchet space \cite[Theorem V.9]{reed_methods_1980}, meaning that it is a Hausdorff space, with the topology induced by the seminorms given by the expectations, and it is complete with respect to those seminorms.
+Physically, this means that measurements on the polynomials (i.
+e., the topology of the seminorms) are enough to identify the states (i.
+e., the singletons, which are closed sets as per Hausdorff).\footnote{This is not true in classical mechanics.
+Even in the single variable case, distinct distributions have the same moments.
+However, if we assume compact support, distinct distributions have distinct moments even in the multi-variable case.\cite{moment_problem_2017} It is curious that the moment problem in quantum mechanics works differently.} Moreover, if we have a sequence for which all expectations converge, we are guaranteed to converge to another state (i.
+e., completeness with respect to the seminorms).
 If any expectation diverges, the limit is not in the space.
-This criteria of convergence seems a lot more physically sound.
+This criterion of convergence seems a lot more physically sound.
 
 To compare Schwartz and Hilbert spaces, the Schwartz space is a subspace of the $L^2$ Hilbert space taken as an inner product space.
-That is, all Schwartz functions are elements of the Hilbert space, they are closed under linear combinations and have the same inner product.
-It is not complete and it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with arbitrary level of precision.
-Or, equivalently, that once a map is defined over the Schwartz space, the map is uniquely extended over the Hilbert space.
-It would seem, then, that all we are losing is the behavior at  actual infinity.
+That is, all Schwartz functions are elements of the Hilbert space; they are closed under linear combinations and have the same inner product.
+It is not complete, and it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with an arbitrary level of precision.
+Or, equivalently, once a map is defined over the Schwartz space, the map is uniquely extended over the Hilbert space.
+It would then seem that all we are losing is the behavior at actual infinity.
 
-The Schwartz space has also a very important property: it is the only space that is closed under Fourier transform.
+The Schwartz space also has a very important property: it is the only space that is closed under the Fourier transform.
 That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property.
@@
 
-If we restrict ourselves to the Schwartz space, then, the only unitary transformations we are interested in are those that map Schwartz function to Schwartz functions.
-These are special cases of unitary transformation over Hilbert spaces, and inherit all their properties, such as the existence of the inverse.
+If we restrict ourselves to the Schwartz space, then the only unitary transformations we are interested in are those that map Schwartz function to Schwartz functions.
+These are special cases of unitary transformation over Hilbert spaces and inherit all their properties, such as the existence of the inverse.
 The change of coordinate we defined in the previous section, then, is automatically ruled out.
-It is interesting to note that two Hilbert spaces connected by such unitary transformations have essentially have two separate theories of distributions as they start from two different sets of objects.
+It is interesting to note that two Hilbert spaces connected by such unitary transformations have essentially two separate theories of distributions as they start from two different sets of objects.
 In our mind, this shows exactly how two inequivalent physical representations can be mathematically equivalent.
-Our suspicion is that the problem of inequivalent mathematical representations in QFT is a consequence of adding another non-physical infinity.
+We suspect that the problem of inequivalent mathematical representations in QFT is a consequence of adding another non-physical infinity.
+%TODO: This last sentence is a bit controversial, and in the main text, it was looser, so I was more comfortable with conjecturing it.
+In the end, I leave it up to you two!
 
-Though we believe the arguments to be thoroughly convincing, we are stopping short from declaring Schwartz spaces physical.
-First, because this is not the goal of the paper, which is just to show the unphysicality of Hilbert spaces.
-Second, because we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The fact that the moment problems works differently in classical and quantum mechanics may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.\footnote{One of the issue is the fact that uniqueness works}
+Though we believe the arguments to be thoroughly convincing, we are stopping short of declaring Schwartz spaces physical.
+First, this is not the paper's goal, which is to show the unphysicality of Hilbert spaces.
+Second, we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The fact that the moment problems work differently in classical and quantum mechanics may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.\footnote{One of the issues is the fact that uniqueness works TODO}
 
-Now, one may be concerned that if we abandon Hilbert spaces we would also have to abandon many convenient tools built on top.
+Now, one may be concerned that if we abandon Hilbert spaces, we would also have to abandon many convenient tools built on top.
 This is not the case.
-When solving an integral, we do sometimes extend the domain for real to complex values, which are unphysical, because complex analysis has nicer mathematical features.
-The result so found, however, is still the valid result on the real domain.
-Similarly, we can pose the problem on the Schwartz space, extend to the Hilbert space or to the space of distributions for calculation, and then bring the result back to the Schwartz space.
-What really need to be rethought, instead, is the construction of state space, especially under infinite operations.
+When solving an integral, we sometimes extend the domain for real to complex values, which are unphysical, because complex analysis has nicer mathematical features.
+The result so found, however, is still a valid result in the real domain.
+Similarly, we can pose the problem on the Schwartz space, extend to the Hilbert space or the space of distributions for calculation, and then bring the result back to the Schwartz space.
+Instead, the construction of state space, especially under infinite operations, needs to be rethought.
 If we want to construct the state space for a composite system of potentially infinitely many subsystems, understanding exactly what infinities are physical and which aren't is exactly the issue.
@@
 
-When mathematicians and philosophers noticed that mathematical theories at the time allowed to express inconsistencies (e.g.
-Russell's paradox, Berry paradox, ...), effort was put into strengthening the foundations of mathematics so that such paradoxes were either resolved or couldn't be expressed.
-With regard to the mathematical foundations of physical theories, we have a similar problem: we can write statements that, though mathematically consistent, are physically inconsistent.
-Looking for these problems and addressing them is ultimately what will lead to a solid foundation, which, we believe, is the prerequisite to solve other outstanding issues in fundamental physics.
+When mathematicians and philosophers noticed that mathematical theories at the time allowed to express inconsistencies (e.
+g., Russell's paradox, Berry paradox, etc.), an effort was put into strengthening the foundations of mathematics so that such paradoxes were either resolved or couldn't be expressed.
+Concerning the mathematical foundations of physical theories, we have a similar problem: we can write statements that, though mathematically consistent, are physically inconsistent.
+Looking for and addressing these problems will ultimately lead to a solid foundation, which, we believe, is the prerequisite to solving other outstanding issues in fundamental physics.
+%TODO: We should discuss the best way to frame this!
 

### 2023-06-27 Articles/2022-ActionPrinciple/ScientificReportsAmended/ActionPrincipleBody.tex
ActionPrinciple: Scientific Reports version

--- 2022-10-13 Articles/2022-ActionPrinciple/ActionPrincipleArchive.tex
@@
-\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
+\documentclass[fleqn,10pt]{wlscirep}
+%\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
 %\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
 %\documentclass[10pt,twocolumn,letterpaper]{article}
+
+\setlength\columnsep{22pt} %To try to make line numbers work in 2-column format
 
@@
 \urlstyle{same}
+\usepackage{lineno}
 \frenchspacing
@@
 \begin{document}
+\linenumbers
 
 \title{Geometric and physical interpretation of the action principle}
-\author{Gabriele Carcassi, Christine A. Aidala}
-\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\author[1]{Gabriele Carcassi}
+\author[1,*]{Christine A. Aidala}
+\affil[1]{Physics Department, University of Michigan, Ann Arbor, Michigan 48109 USA}
+
+\affil[*]{caidala@umich.edu}
+
+%\author{Gabriele Carcassi, Christine A. Aidala}
+%\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
 
@@
 
-\section{Introduction}
+\section*{Introduction}
 
-While the principle of stationary action is regarded by many as one of the most important tools in physics,\cite{hamilton1834general, heaviside1903, feynman1942principle, brenier1989least, rojo2018principle} its physical meaning is not completely clear.
-First of all, the typical characterization of the Lagrangian as the difference between kinetic and potential energy fails even for simple systems, like a charged particle under a magnetic field.
+While the principle of stationary action is regarded by many as one of the most important tools in physics, its physical meaning is not completely clear.\cite{hamilton1834general, heaviside1903, feynman1942principle, brenier1989least, rojo2018principle} First of all, the typical characterization of the Lagrangian as the difference between kinetic and potential energy fails even for simple systems, like a charged particle under a magnetic field.
 Moreover, the Lagrangian for a system is not uniquely defined, making the actual value of the action for a path not directly physically significant.
@@
 The two aspects that have been sorely lacking, and that we provide, are (1) a clear geometric interpretation of the action principle and (2) a tight connection between the math and the physics it represents.
-To make the result accessible to the widest audience, we first cover the case of a single degree of freedom using standard vector calculus.
+[This has been the hardest problem to solve, as the modern mathematical concepts and notation have departed from physical intuition.
+We will therefore use, instead of the standard notation of differential geometry, an extended version of the one physicists use in general relativity.] To make the result accessible to the widest audience, we first cover the case of a single degree of freedom using standard vector calculus.
 We then proceed to the case of multiple independent degrees of freedom using tools from differential geometry.
-The main article will presents all the key point needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the appendix.
+The main article will presents all the key points needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the Supplementary Information.
 
-\section{One degree of freedom}
+\section*{One degree of freedom}
 
@@
 \end{equation}
-that describes how states move in time.\footnote{Where possible, we will be writing the same expression in both vector calculus and component notations.} The time component of the displacement vector field is constrained, as we have
+that describes how states move in time.
+[Where possible, we will be writing the same expression in both vector calculus and component notations.] In dynamical system literature, this is referred to as the vector field of the dynamical system.
+The time component of the displacement vector field is constrained, as we have
 \begin{equation}\label{sdof_time_constraint}
@@
 Alternatively, if we assign a probability, or probability density, to each trajectory, the assumptions requires that probability not to change, so integrating the probability over a closed surface must yield zero.
-However we see it, assumption \ref{assum_detrev} means the field is divergence-free,\footnote{Given that this is a three dimensional space, we can use the standard tools of vector calculus.} that is
+However we see it, assumption \ref{assum_detrev} means the field is divergence-free.
+[Given that this is a three dimensional space, we can use the standard tools of vector calculus.] That is,
 \begin{equation}\label{sdof_div_free}
@@
 \end{equation}
-where $H$ is a suitable function of $q$, $p$ and $t$. By applying definition \ref{sdof_displacement} and expanding \ref{sdof_displacement_potential} with \ref{sdof_potential_expression}, we have
+where $H$ is a suitable function of $q$, $p$ and $t$.
+The potential $\vec{\theta}$ is closely related to the canonical one-form of symplectic geometry and the contact form of contact geometry.
+By applying definition \ref{sdof_displacement} and expanding \ref{sdof_displacement_potential} with \ref{sdof_potential_expression}, we have
 \begin{equation}\label{sdof_Ham_eq}
@@
 	- \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} &= \iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
-	=  \oint_{\partial \Sigma = \gamma \cup \gamma'} \vec{\theta}  \cdot d\vec{\gamma} 
+	&=  \oint_{\partial \Sigma = \gamma \cup \gamma'} \vec{\theta}  \cdot d\vec{\gamma}  \\
+&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma}' \\
 	&= \delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}.
@@
 
-\section{Multiple degrees of freedom}
+\section*{Multiple degrees of freedom}
 
@@
 \end{equation}
-which we call the state counting form.\footnote{Note that this approach is consistent with statistical mechanics\cite{peliti2011statistical}.}
+which we call the state counting form. [Note that this approach is consistent with statistical mechanics\cite{peliti2011statistical}.]
 
@@
 as the parallelogram identified by $\vec{v}$ and $\vec{w}$ in that order will be the same as the one identified by $\vec{w}$ and $\vec{v}$ with opposite orientation.
-The form $\omega$ will also be closed, meaning\footnote{In our notation, $d\Sigma$ represents the infinitesimal surface element of integration, which is the argument of the form.}
+The form $\omega$ will also be closed, meaning
 \begin{equation}\label{mdof_closed_form}
@@
 over all closed contractible surfaces.
-This stems from \ref{assum_indep}.
+[In our notation, $d\Sigma$ represents the infinitesimal surface element of integration, which is the argument of the form.] This stems from \ref{assum_indep}.
 Suppose you take a surface and translate it along another independent variable.
@@
 Therefore if we construct any parallelepiped at equal time, two opposite sides will contain the same number of states with opposite orientation, and therefore the integral of $\omega$ over the whole surface will be zero.
+Those familiar with differential geometry will recognize $\omega$ as the symplectic form of symplectic manifolds and of the symplectic bundles of contact manifolds.
 
-Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative\footnote{We write the exterior derivative as $\partial \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.} of a covector $\theta$:
+Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative of a covector $\theta$:
 \begin{equation}\label{mdof_form_potential}
@@
 		\theta &= \theta_a e^a = \theta_{q^i} e^{q^i} + \theta_{p_i} e^{p_i} + \theta_t e^t \\
-		\omega &= - \partial \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
+		\omega &= - \partial \wedge \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
 	\end{aligned}
@@
 The minus sign is required to match conventions.
+[We write the exterior derivative as $\partial \wedge \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.]
 
@@
 \begin{equation}\label{mdof_form_components}
-	\omega_{ab} = (-\partial\theta)_{ab} = \begin{bmatrix}
+	\omega_{ab} = (-\partial\wedge\theta)_{ab} = \begin{bmatrix}
 		0 & \delta^j_i & \partial_{q^i} H \\
@@
 
-\section{Discussion and conclusions}
+\section*{Discussion and conclusions}
 
@@
 In this case the ``Lagrangian'' $L(q,\dot{q},p,t)$ depends on conjugate momentum as well.
-Also note that the cases where the Lagrangian does not admit a Hamiltonian are exactly the cases where the action principle fails to yield a unique solution.
+That is, the principle holds on all Hamiltonian systems, whether or not they admit a Lagrangian.
+Conversely, the cases where the Lagrangian does not admit a Hamiltonian are exactly the cases where the principle of stationary action fails to yield a unique solution, thus does not hold.
+These are the cases where the Lagrangian is not hyperregular.
+Therefore the principle of stationary action holds for all Lagrangian systems that admit a Hamiltonian, but also for all Hamiltonian systems whether of not they admit a Lagrangian.
 In this sense, the action principle is better understood as a feature of Hamiltonian mechanics, instead of Lagrangian mechanics\cite{souriau1970structure, arnold1989mathematical}.
@@
 
+\vspace{-1mm}
+\section*{Data availability}
+All data generated or analysed during this study are included in this published article and its supplementary information files.
+
+\vspace{-1mm}
+\section*{Acknowledgments}
+This paper forms part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which
+the basic laws can be rigorously derived. The authors would like to thank Alejandro Uribe-Ahumada and Daniel Burns for helpful discussions.
+
 \bibliography{bibliography}
 
-\section*{Appendix}
-
-In this section we will review the arguments in more detail.
-Given that this article is still aimed at a physics audience, the focus will be conceptual rigor, rather than mathematical rigor.
-What is important is to establish a precise map between physical concepts and their mathematical representation, so that it is clear what physical requirement is captured by each mathematical definition.
-
-We divide the appendix into three sections: the first will address the single d.o.f.
-case, the second will give a geometric interpretation of the tools of differential geometry that is suitable for physicists, and lastly we address the multiple d.o.f.
-case.
-
-\subsection*{One degree of freedom}
-
-For the case of one degree of freedom we are going to use vector calculus. This is possible since the problem can be studied with three variables (position, momentum and time)
-\begin{equation}
-\tag*{(\ref{sdof_variables})}
-	\xi^a = [ q, p, t]
-\end{equation}
-which is the only setting where vector calculus truly works.
-In this sense, nature gave us a hand as three dimensional spaces are much more intuitive for us, and we can therefore develop the ideas in a familiar context before generalizing.
-
-We break down the problem into sequential steps, which serves as a summary for the whole argument. We will have to:
-\begin{enumerate}[label=(\roman*)]
-	\item show that assuming determinism and reversibility \ref{assum_detrev} is equivalent to assuming the existence of the displacement field $\vec{S}$ defined in \ref{sdof_displacement} and its being divergence-free as in \ref{sdof_div_free}
-	\item show that the vector potential $\vec{\theta}$ for $\vec{S}$ defined in \ref{sdof_displacement_potential} can be expressed, without loss of generality, as \ref{sdof_potential_expression}
-	\item show that \ref{sdof_div_free} and \ref{sdof_potential_expression} recover Hamilton's equations as in \ref{sdof_Ham_eq}
-	\item show that the flow of $\vec{S}$ through a surface $\Sigma$ delimited by a path $\gamma$ and its variation $\gamma'$ is the variation of the integral of the vector potential, as claimed in \ref{sdof_action}
-	\item show that the integral of the vector potential is stationary for and only for the solutions of the equations of motion
-	\item show that the integral of the vector potential reduces to the action if and only if \ref{assum_kineq} applies.
-\end{enumerate}
-
-For (i), determinism and reversibility means that giving the state at some point in time means constraining the state at all times, that the system can evolve in only one way.
-In math terms, if $X$ is the manifold of all possible states at all times, given $P = \{q_0, p_0, t_0\}$ there will be only one possible evolution $\gamma : \mathbb{R} \to X$ such that $\gamma(t_0) = P$.
-
-If $X$ were a discrete space, the previous mathematical condition would be sufficient, but on a manifold it is not.
-Consider a damped harmonic oscillator: it satisfies the previous condition, yet the system will concentrate at the equilibrium.
-What needs to happen is that densities also need to be mapped exactly.
-That is, if a certain fraction of the system is in a given initial state, the same fraction of the system must be in the corresponding final state.
-Equivalently, if the system has a given probability of starting in a given initial state, it must have the same probability of ending in the corresponding final state.
-In mathematical terms, if $\rho : X \to \mathbb{R}$ is a statistical distribution (or a probability density) defined on our extended phase space and $\gamma$ is a possible evolution, we must have $\rho(\gamma(t_0)) = \rho(\gamma(t_1))$.
-
-This requirement tells us two things.
-First, that $X$ must be a differentiable manifold.
-To define densities over a set of variables and map them backward and forward in time Jacobians must be defined, which means the derivative across state variables must be defined as well, in particular with respect to time, which plays the double role of a coordinate of $X$ and the evolution parameter.
-This means that the evolutions must be differentiable curves and therefore the displacement $\vec{S}$ must be defined, which leads to
-\begin{equation}
-\tag*{(\ref{sdof_displacement})}
-	\begin{aligned}
-		\vec{S} &= \left[ \frac{dq}{dt},\frac{dp}{dt},\frac{dt}{dt} \right] \\
-		&= S^a e_a = \frac{d\xi^a}{dt} e_a .
-	\end{aligned}
-\end{equation}
-The double role of time gives us
-\begin{equation}
-\tag*{(\ref{sdof_time_constraint})}
- 	S^t=\frac{dt}{dt}=1.
- \end{equation}
-
-Secondly, since the densities must be constant over evolutions, the flow of these densities through a closed surface in $X$ must be zero: each evolution will enter and exit the region delimited by the surface, giving no net contribution.
-Mathematically this requires the displacement to be divergence-free, which leads to
-\begin{equation}
-\tag*{(\ref{sdof_div_free})}
-	\nabla \cdot \vec{S} = \partial_a S^a = 0.
-\end{equation}
-
-For (ii), since $\vec{S}$ is divergence-free, we can find a vector potential
-\begin{equation}
-\tag*{(\ref{sdof_displacement_potential})}
-	\begin{aligned}
-		\vec{\theta} &= [\theta_q, \theta_p, \theta_t] \\
-		&= \theta_a e^a \\
-		\vec{S} &= - \nabla \times \vec{\theta} \\
-		&= - \epsilon^{abc} \partial_b \theta_c \, e_a. \\
-	\end{aligned}
-\end{equation}
-Note that the vector potential is not uniquely identified, since $\nabla \times(\vec{\theta} + \nabla f) = \nabla \times \vec{\theta}$ and therefore the displacement field $\vec{S}$ remains unchanged.
-This is commonly referred to as gauge freedom in physics, and it is analogous to what happens for a magnetic field or for an incompressible fluid.
-
-We use this gauge freedom to set the momentum component to zero.
-In fact, we can always choose $f(q,p,t)$ such that $\partial_p f = -\theta_p$ and redefine $\vec{\theta}$ to be $\vec{\theta} + \nabla f$.
-Therefore we can set
-\begin{equation}
-	\vec{\theta} = [\theta_q, 0, \theta_t]
-\end{equation}
-without loss of generality.
-
-We now use constraint \ref{sdof_time_constraint} on the time component. We have
-\begin{align}
-	S^t = \frac{dt}{dt} &= - \left(\frac{\partial}{\partial q}  \theta_p - \frac{\partial}{\partial p}  \theta_q\right) \\
-	&= - \left(\frac{\partial}{\partial q}  0 - \frac{\partial}{\partial p}  \theta_q\right) \\
-	& = \frac{\partial \theta_q}{\partial p} = 1
-\end{align}
-Integrating, we have $\theta_q = p + g(q,t)$ where $g(q,t)$ is an arbitrary function.
-This function can be set to zero without loss of generality as it can be removed with a gauge transformation where $f(q,t)$ does not depend on $p$, and therefore $\theta_p$ will remain unchanged.
-Therefore we have:
-\begin{equation}
-	\vec{\theta} = [p, 0, \theta_t]
-\end{equation}
-Lastly, we rename the last component as $\theta_t = -H$ which leads to 
-\begin{equation}
-\tag*{(\ref{sdof_potential_expression})}
-	\begin{aligned}
-		\vec{\theta} &= [p, 0, -H] \\
-		&= p e^q - H e^t.
-	\end{aligned}
-\end{equation}
-
-For (iii), we expand each component of \ref{sdof_displacement_potential} using \ref{sdof_potential_expression}.
-Again note the double role of time: $\frac{dp}{dt}$ is not the same as $\frac{\partial p}{\partial t}$.\footnote{We could be more precise and use $\tau$ for the evolution parameter, leaving $t$ just for the coordinate.
-The treatment would automatically acquire a relativistic flavor and would enlighten us to other interesting results, though it would distract from the main focus of this paper, the action principle.
-} In the first case we are taking a total derivative along the evolution, and therefore the momentum can change.
-In the second case we are taking a partial derivative at constant $q$ and $p$ by definition, and therefore $\frac{\partial p}{\partial t}=0$.
-We have
-\begin{equation}
-\begin{aligned}
-	S^q = \frac{dq}{dt}
-	&= - \left( \frac{\partial}{\partial p} \theta_t - \frac{\partial}{\partial t} \theta_p \right) \\
-	&= - \left( \frac{\partial}{\partial p} (-H) - \frac{\partial}{\partial t} 0 \right) \\
-	& = \frac{\partial H}{\partial p}
-\end{aligned}
-\end{equation}
-\begin{equation}
-\begin{aligned}
-	S^p = \frac{dp}{dt}
-	&= - \left( \frac{\partial}{\partial t} \theta_q - \frac{\partial}{\partial q} \theta_t \right) \\
-	&= - \left( \frac{\partial}{\partial t} p - \frac{\partial}{\partial q} (-H) \right) \\
-	& = - \frac{\partial H}{\partial q}
-\end{aligned}
-\end{equation}
-\begin{equation}
-\begin{aligned}
-	S^t = \frac{dt}{dt}
-	&= - \left( \frac{\partial}{\partial q} \theta_p - \frac{\partial}{\partial p} \theta_q \right) \\
-	&= - \left( \frac{\partial}{\partial q} 0 - \frac{\partial}{\partial p} p \right) \\
-	& = 1
-\end{aligned}
-\end{equation}
-We have recovered Hamilton's equations
-\begin{equation}
-\tag*{(\ref{sdof_Ham_eq})}
-\begin{aligned}
-	\frac{dq}{dt} &= \frac{\partial H}{\partial p} \\
-	\frac{dp}{dt} &=-\frac{\partial H}{\partial q}
-\end{aligned}
-\end{equation}
-as the equations for a deterministic and reversible system.
-The only thing that the equations say is that states move in time at the same rate (i.e.
-$\frac{dt}{dt} = 1$) with an incompressible flow (i.e.
-no states are created or destroyed).
-That is the whole physical and geometric content of those equations.
-
-Note that if we started from Hamilton's equations, the system would be deterministic and reversible: position and momentum at a fixed time would indeed allow us to pick a single differentiable evolution and, by Liouville's theorem, densities would be constant along such evolution.
-Therefore we have an equivalence between assumption \ref{assum_detrev} and Hamilton's equations \ref{sdof_Ham_eq} in the single degree of freedom case.
-
-Now that we have developed a clear geometric and physical understanding of Hamilton's equations, we can concentrate on the action principle.
-For (iv) we can apply Stokes' theorem.
-We take a path $\gamma$ and a variation $\gamma'$.
-Since they share the endpoints, taken together they form a closed curve where $\gamma'$ is taken with the opposite orientation.
-Because $\vec{S}$ is divergence-free, the flow of $\vec{S}$ through a surface depends only on its contour.
-Moreover, by Stokes' theorem, the flow through the surface will be equal to the integral of the potential over the contour.
-Putting it all together we have
-\begin{equation}
-\tag*{(\ref{sdof_action})}
-\begin{aligned}
-	- \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} &=
-	\iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
-	&= \oint_{\partial \Sigma} \vec{\theta}  \cdot d\vec{\gamma} \\
-	&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma}' \\
-	&= \delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}.
-\end{aligned}
-\end{equation}
-
-For (v), note that in the neighborhood of $\gamma$ the infinitesimal surface element $d\vec{\Sigma}$ has one side along $\gamma$ itself.
-Therefore we can write $d\vec{\Sigma} = d\vec{\gamma} \times d\vec{\lambda}$, where $d\vec{\lambda}$ is the infinitesimal displacement that connects $\gamma$ with $\gamma'$.
-The integrand then becomes the triple product
-\begin{equation}
-	\vec{S} \cdot d\vec{\gamma} \times d\vec{\lambda}.
-\end{equation}
-We can understand it geometrically as the volume element given by the displacement $\vec{S}$, the infinitesimal line element $d\vec{\gamma}$ and the infinitesimal variation $d\vec{\lambda}$ that moves $\gamma$ to $\gamma'$.
-
-If $\gamma$ is a solution of the equations of motion, $\gamma$ is a field line of $\vec{S}$: $\gamma$ is tangent to $\vec{S}$ at all points.
-We can write $d\vec{\gamma} = \vec{S} dt$.
-The triple product, in this case, is zero regardless of what $d\vec{\lambda}$ is.
-We can reason the other way: we look for a line $\gamma$ such that the triple product is zero regardless of what $d\vec{\lambda}$ is.
-The only way for that to happen is if $\gamma$ is tangent to $\vec{S}$ at each point, which means $\gamma$ is a solution of the equations of motion.
-This means that the paths that make the line integral of the vector potential stationary are those and only those that solve the equations of motion.
-This recovers the principle of stationary action in Hamiltonian form
-\begin{equation}
-\tag*{(\ref{sdof_stationary_action})}
-	0 =\delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} = - \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} 
-\end{equation}
-
-We want to stress that this variational principle applies to any divergence-free field. For example, for a magnetic field we could write:
-\begin{equation}
-	\iint_{\Sigma} \vec{B} \cdot d\vec{\Sigma} = \delta \int_{\gamma} \vec{A} \cdot d\vec{\gamma}.
-\end{equation}
-We would find that the field lines of $\vec{B}$ are the lines and only the lines for which the integral of the vector potential is stationary.
-Similarly, we would find that the streamlines of an incompressible fluids are the lines and only the lines for which the integral of the vector potential of the flow is stationary.
-While these properties may be of limited practical use, the geometric insight is the same.
-
-For (vi), if the assumption \ref{assum_kineq} of kinematic equivalence  holds, we can from a trajectory in space $q(t)$ reconstruct the state at any time.
-That is, if we are given $q(t)$ we can reconstruct the full evolution $\gamma(t)$ in the extended phase space.
-
-We saw before that the evolution must be differentiable, which means $q(t)$ must also be differentiable.
-Therefore the velocity $\dot{q} = \frac{dq}{dt}$ is defined at every instant.
-Using assumption \ref{assum_detrev}, we know that each evolution is fully identified by position and momentum at a given time.
-Therefore the space of all possible evolutions is two-dimensional.
-Given that for each evolution $\gamma(t)$ there is one and only one spatial trajectory $q(t)$, the space of all spatial trajectories must also be two-dimensional and therefore fully identified by position and velocity at a given instant.
-Therefore there must be a map $p(q, \dot{q}, t)$ that allows us to reconstruct the momentum at a given time by knowing position and velocity at that time.
-
-We can use this map to express the Hamiltonian form of the principle of stationary action only in terms of position and velocity. We have
-\begin{equation}
-\tag*{(\ref{sdof_Lagrangian})}
-	\begin{aligned}
-		\delta \int_\gamma \vec{\theta} \cdot d\vec{\gamma} &= \delta \int^{t_2}_{t_1} \vec{\theta} \cdot \frac{d\vec{\gamma}}{dt} dt \\
-		&= \delta \int^{t_2}_{t_1} \left(\theta_q \frac{dq}{dt} + \theta_p \frac{dp}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
-		&= \delta\int^{t_2}_{t_1} \left(p \frac{dq}{dt} + 0 \frac{dp}{dt} - H \frac{dt}{dt}\right) dt \\
-		&= \delta \int^{t_2}_{t_1} \left(p \frac{dq}{dt} - H\right) dt \\
-		&= \delta \int^{t_2}_{t_1} L(q, \dot{q}, t) \, dt = 0
-	\end{aligned}
-\end{equation}
-We recognize the principle of stationary action in standard form.
-
-Note that the argument can proceed in the opposite way.
-If we have a system for which the principle of stationary action yields a unique solution, then we can write $p=\frac{\partial L}{\partial \dot{q}} = p(q, \dot{q}, t)$ which means we are able to reconstruct the momentum, and therefore the state, at each time.
-The system satisfies \ref{assum_kineq}.
-Moreover, we can find a unique Hamiltonian such that the solutions of the equations of motion are exactly those paths that satisfy the principle of stationary action.
-Therefore the system satisfies \ref{assum_detrev} as well.
-
-Before proceeding to the generalization, a few observations are in order.
-\begin{itemize}
-	\item We stress that the vector potential $\vec{\theta}$ is not uniquely defined, and therefore neither are the action nor the Lagrangian.
-Giving precise physical characterizations of those two objects, then, is in our view a hopeless task.
-	\item We found that the Hamiltonian $H$ is the time component of the vector potential $\vec{\theta}$, not a scalar. This relativistic aspect emerges naturally out of the assumptions.
-	\item One can, in principle, write Lagrangian systems for which the action principle does not yield a single solution (e.g.
-for $L=0$ all paths are stationary).
-Consistently, assumption \ref{assum_detrev} fails.
-\end{itemize}
-
-\subsection*{Introduction to differential topology}
-
-The generalization to multiple d.o.f will require the use of differential topology.
-These tools are not typically taught in a standard physics curriculum.
-If they are, their presentation is dense and abstract.
-Apart from the abstract definitions, the notation gets in the way of the physics.
-For example, writing $\frac{\partial}{\partial x^i}$ for the vector basis makes sense if one is studying the space of linear operators defined by derivations.
-When studying the space of velocities, however, writing $\vec{v} = v^i\frac{\partial}{\partial x^i}$ decreases understanding.
-Therefore we will use a notation and nomenclature that is closer to what physicists already use (e.g.
-Einstein notation) and need to picture.
-For example, we will use $e_i$ for vector basis and $e^i$ for covector (dual) basis.
-To those that believe we should just use the ``correct'' mathematical nomenclature, we retort that there is no such thing.
-Notations and names are chosen within a certain context to make the points of the discourse more clear.
-
-What we will give here, then, is not a fully rigorous introduction to the subject, but rather a geometric understanding of what these tools describe and a physical motivation as to why they are useful.
-We hope this will make it easier to follow the next section and to build a geometric and physical intuition of the principle of stationary action, which is the focus of this article.
-
-%This very result, and other connected to it, became apparent to us precisely because we started to translate the language of differential geometry into a form that is less abstract.
-These results have been in front of our known noses, hidden in plain sight by a notation, useful to some mathematicians, but arcane to others.
-While we don't claim our notation is ``the best'', it is better for our purposes than the standard ones.
-We understand that there is a cost associated to learning a new notation, but to us that cost is essentially learning exactly why the tools are useful and what they represent in physics.
-It is worth the long term investment
-
-%In our opinion, , and In order to fully understand the geometric and physical meaning of the general case, it helps to have a crisp intuition for the tools of differential topology.
-We will therefore go through the basic motivation of differential topology from a physicist perspective, so that we can see what differential forms and exterior derivatives are capturing at a a physical level.
-
-The core objective of using differential topology in physics is to characterize those quantities that are associated with a region of space and can be understood as sums of independent contributions.
-The mass $m$ within a region $V$ can be understood as the sum of the contributions of a mass density $\rho$ within each infinitesimal volume $dV$:
-\begin{equation*}
-	m(V) = \iiint_V \rho dV.
-\end{equation*}
-The magnetic flux $\Phi$ through a surface $\Sigma$ can be understood as the sum of the contributions of the magnetic field $\vec{B}$ over each infinitesimal $\vec{d\Sigma}$:
-\begin{equation*}
-	\Phi(\Sigma) = \iint_\Sigma \vec{B} \cdot \vec{d\Sigma}.
-\end{equation*}
-The work $W$ over a path $\lambda$ can be understood as the sum of the contributions of a force field $\vec{f}$ over each infinitesimal line $\vec{d\gamma}$:
-\begin{equation*}
-	W(\gamma) = \int_\gamma \vec{f} \cdot \vec{d\gamma}.
-\end{equation*}
-
-Though the dimensionality changes, a pattern emerges: the functionals $W$, $\Phi$ and $m$ all take a region of space while $\vec{f}$, $\vec{B}$ and $\rho$ act on infinitesimal regions.
-However, the vector representation is in terms of different objects (i.e.
-vector, pseudo-vector and pseudo-scalar) and operations.
-Thus it fails to provide a nice generalization.
-
-Suppose we wrote
-\begin{equation*}
-\begin{aligned}
-	W(\gamma) &= \int_\gamma f(d\gamma) \\
-	\Phi(\Sigma) &= \iint_\Sigma B(d\Sigma) \\
-	m(V) &= \iiint_V \rho(dV).
-\end{aligned}
-\end{equation*}
-Here we don't have vectors, but functions of infinitesimal regions, which we call differential forms.
-The force, in this notation, is a one-form (or covector) as it takes one dimensional infinitesimal regions (i.e.
-vectors); the magnetic field is a two-form; the density is a three-form.
-We can also say that a scalar field, like the temperature, is a zero-form.
-
-An infinitesimal region can also be understood as a parallelepiped, which is fully identified by its sides.
-Therefore a differential form can be understood as acting on a set of vectors, the sides of the parallelepiped, that matches the dimensionality of the form.
-A one-form will take one vector, a two-form two vectors and so on.
-All forms must be anti-symmetric because switching the order of the sides does not change the parallelepiped, just its orientation.
-
-This picture is much easier to generalize to any dimensionality.
-If we note $S^k$ the space of all the $k$ dimensional subregions, a $k$-functional $F : S^k \to \mathbb{R}$ is a linear functional that takes a $k$-surface $\sigma$ and returns a number.
-This can be expressed as
-\begin{equation*}
-	F(\sigma) = \int_\sigma \omega(d\sigma)
-\end{equation*}
-the integral of a $k$-form $\omega : V^k \to \mathbb{R}$ that takes an infinitesimal $k$-surface $d\sigma$ (or equivalently a set of $k$ vectors) and returns a number.
-
-This allows us to identify properties and theorems that apply to all functionals and forms regardless of their dimensionality.
-For example, if $\sigma_1$ and $\sigma_2$ are two disjoint regions, $F(\sigma_1 \cup \sigma_2) = F(\sigma_1) + F(\sigma_2)$, which also tells us $F(\emptyset) = 0$ for any functional.
-Additionally, $k$-forms are linear maps and can be written in terms of the tensor product of the co-basis $e^a$.
-We have:
-\begin{equation*}
-	\begin{aligned}
-		f &= f_a e^a \\
-		B &= B_{ab} e^a \otimes e^b \\
-		\rho &= \rho_{abc} e^a \otimes e^b \otimes e^c.
-	\end{aligned}
-\end{equation*}
-These tensors are all anti-symmetric as the forms are anti-symmetric.
-
-%If $F_k$ is the set of all linear $k$-functionals and $\Omega_k$ the set of all linear $k$-forms, we can understand $\Omega_k$ as the infiniteimsal verson of $F_k$.
-
-%of one, two or three dimensions
-
-%We say that $m$ is a functional that takes a three dimensional region and returns a number.
-This functional is linear, in the sense that it respects the limit and it is additive over disjoint regions.
-That is, $m(V_1 \cup V_2) = m(V_1) + m(V_2)$ if $V_1 \cup V_2 = \emptyset$.
-
-%Since $k$-form are also linear maps, they form a vector space and can be written in terms of a basis.
-In the same way we have a basis $e_a$ for vectors, we will have a basis $e^a$ for one-forms (or covectors) such that $e^a(e_b) = \delta^a_b$.
-We have:
-%Additionally, all forms are anti-symmetric in all indexes: a changing the order of the vectors reduces to a change of the orientation.
-Mathematically, this is why one introduces the wedge product $\wedge$ which is an anti-symmetrized version of the tensor product.
-Given that physicists are already familiar with the idea of anti-symmetric tensors, we think it's better to not introduce the wedge product and work in terms of tensors.
-
-We can see how this works for a force applied along a path. The infinitesimal displacement along $\gamma$ is given by
-\begin{equation*}
-	d\gamma = dx^a e_a.
-\end{equation*}
-Therefore 
-\begin{equation*}
-	\begin{aligned}
-		W(\gamma) &= \int_\gamma f(d\gamma) = \int_\gamma  f_a e^a(dx^b e_b) \\
-		&= \int_\gamma  f_a dx^b e^a( e_b)
-		= \int_\gamma  f_a dx^b \delta^a_b
-		= \int_\gamma  f_a dx^a.
-	\end{aligned}
-\end{equation*}
-
-Of all generalized properties, the most impressive one is certainly the generalized Stokes' theorem. Compare the divergence (or Gauss's) theorem:
-\begin{equation*}
-	\iiint_V (\nabla \cdot \vec{B}) \, dV = \oiint_{\Sigma = \partial V} \vec{B} \cdot \vec{d\Sigma}
-\end{equation*}
-with Stokes' theorem
-\begin{equation*}
-	\iint_\Sigma (\nabla \times \vec{f}) \cdot \vec{d\Sigma} = \oint_{\gamma = \partial \Sigma} \vec{f} \cdot \vec{d\gamma}
-\end{equation*}
-and the divergence theorem
-\begin{equation*}
-	\int_\gamma (\nabla \varphi) \cdot \vec{d\gamma} = \left[ \varphi \right]_A^B .
-\end{equation*}
-In all three cases we have a region where we define the integral of some differential operator applied to a field.
-This is equated to the integral over the boundary of the field itself.
-But again, the vector calculus notation prevents us from generalizing as the differential operator is somewhat different in each of the three cases.
-
-To understand conceptually that a generalization must exist, suppose we have a $k$-functional $F$ and its corresponding $k$-form $\omega$.
-By definition $F$ acts on $k$-surfaces.
-Now suppose we have a $k+1$-dimensional surface $\Sigma$.
-This is not a valid argument for $F$.
-However, its boundary $\partial \Sigma$ will be a $k$-dimensional surface, which is a valid argument for $F$.
-Therefore from any $F$ we can create a $k+1$-functional, which we can call the exterior functional of $F$, noted $\partial F$, such that $\partial F(\Sigma) = F(\partial \Sigma)$.
-If we look at all three cases above, that is exactly what is happening: the gradient, the curl and the divergence lead to a functional one dimension higher by applying the original operation on the boundary.
-
-Now, since every functional has a corresponding form, $\partial F$ must have a corresponding $k+1$-form which must be fully identified by $\omega$. We note it as $\partial \omega$ and write
-\begin{equation*}
-	\partial F(\Sigma) = \int_\Sigma \partial \omega(d\Sigma) = \oint_{\sigma = \partial \Sigma} \omega(d\sigma) = F(\partial \Sigma).
-\end{equation*}
-This is the generalized Stokes' theorem and $\partial \omega$ is the exterior derivative of $\omega$. We use $\partial$ as we leave $d$ reserved for the infinitesimal regions.
-
-With a few steps, one can see that the divergence, curl and gradient correspond to the exterior derivative in three dimensions for a zero-form, one-form and two-form respectively, and that the three theorems of vector calculus are particular instances of this generalized one.
-
-Moving on, the identities
-\begin{equation*}
-	\begin{aligned}
-		\nabla \cdot \nabla \times \vec{f} &= 0 \\
-		\nabla \times \nabla \varphi &= 0
-	\end{aligned}
-\end{equation*}
-both correspond to the single identity
-\begin{equation*}
-	\partial \partial \omega = 0,
-\end{equation*}
-which states that the exterior derivative applied twice is also zero.
-This identity can be understood in terms of the functionals.
-We have  $\partial \partial F(\Sigma) = F(\partial \partial \Sigma)$ and the boundary of a boundary $\partial \partial \Sigma = \emptyset$ is always the empty set.
-So, no matter what $F$ or $\Sigma$ is, we are evaluating a functional on an empty set, which is zero.
-
-In the same way that a divergence-free field can be written as the curl of a vector potential, or a curl-free field can be written as the gradient of a scalar potential, a form whose exterior derivative is zero $\partial \omega = 0$ (a closed form), can be written as the exterior derivative $\omega = \partial \theta$ of a lower dimensional form, at least in some neighborhood.
-In terms of functionals, this corresponds to the case where $F(\sigma)=0$ for all contractible surfaces, those surfaces that can be morphed to a point with a continuous transformation.
-If $F(\sigma)=0$ over all surfaces, instead, then the corresponding form is exact.
-
-The overall point here is that understanding differential topology in terms of functionals of regions gives us much better motivation and intuition for the math.
-
-\iffalse
-
-%If we compute the exterior derivative for zero, one and two-form
-
-%The last element we need to introduce is the exterior derivative.
-We are actually much better equipped to understand this operation as it is much more intuitive on the linear functionals than the corresponding forms.
-
-%First, note that every $k$-surface has a $k-1$ boundary.
-For example, a three dimensional volume would be enclosed by a two dimensional surface.
-Therefore we have a boundary operator $\partial \sigma^k \in S^{k-1}$.
-If the surface does not have a boundary, like the surface of a sphere, then the boundary is the empty set $\emptyset$.
-Note that the all boundaries are closed surfaces.
-Therefore the boundary of a boundary is the empty set.
-That is $\partial \partial \sigma = 0$ for all $\sigma$.
-
-Now recall Stokes' theorem:
-On the left side we have a two-functional, a functional of a surface.
-On the right side we have a one-functional, a functional of a line.
-The two-functional can be defined in terms of the one-functional: given a surface, take its boundary and pass it to the one functional.
-Recall Gauss's theorem:
-Something similar is happening except we have a three functional on the left and a two functional on the right. Recall the gradient theorem:
-\begin{equation*}
-	\int_\gamma \nabla \varphi \cdot d\gamma = \left[ \varphi \right]_A^B
-\end{equation*}
-Here we have a one functional on the left and function of points on the right, which can be understood as a zero functional.
-We note a pattern, but again the vector calculus notation does not help us express the pattern.
-
-Differential topology generalizes the pattern through the generalized Stokes' theorem
-\begin{equation*}
-	\int_\sigma \partial \omega(d\sigma) = \oint_{\partial \sigma} \omega(d\partial\sigma)
-\end{equation*}
-Our notation differs from the standard in that we use $\partial$ as the symbol for the exterior derivative and we make explicit the argument of the forms.
-We still have, however, to understand what is it that we are generalizing.
-This is better understood in terms of linear functionals.
-
-Suppose we have a $k$-functional $F$, for example work along a path.
-Each $k$ dimensional surface $\sigma$ will have a boundary $\partial \sigma$ that is one dimensional.
-In the example, a two dimensional surface will identify a line.
-This means that, to each $k$-dimensional surface $\sigma$ we can associate a number $F(\partial \sigma)$.
-In the example, to a two-dimensional surface we can associate the work along its boundary.
-Proceeding this way we can define the exterior functional, which we note $\partial F$ and note that $\partial F(\sigma) = F(\partial \sigma)$.
-The notation is admittedly a bit unusual, but it allows us to show very easily that the exterior functional of the exterior functional is always zero.
-In fact
-\begin{equation*}
-	\partial \partial F(\sigma) = \partial F(\partial \sigma) = F(\partial \partial \sigma) = F(\emptyset) = 0
-\end{equation*}
-This is a direct consequence of the boundaries not having boundaries.
-
-Since every $k$-functional has a corresponding $k$-form, so will the exterior functionals.
-Using Stokes' theorem we find that the form associated to the exterior functional is exactly the exterior derivative.
-
-Obviously a full treatment of all these concepts goes beyond the scope of this appendix.
-Yet, we believe it is important to see that we can have a good understanding of these tools in terms of ideas that can be meaningful to physicists.
-
-\fi
-
-\subsection*{Multiple degrees of freedom}
-
-The case of multiple degrees of freedom proceeds in the same way as the single degree of freedom.
-The main difference is that we will work with a system described by $N$ degrees of freedom, which is therefore described by the $2N+1$ variables
-\begin{equation}
-\tag*{(\ref{mdof_variables})}
-	\xi^a = [ q^i, p_i, t].
-\end{equation}
-We will have to:
-\begin{enumerate}[label=(\roman*)]
-	\item show that assuming determinism and reversibility \ref{assum_detrev} and independence of degrees of freedom \ref{assum_indep} is equivalent to assuming the existence of the displacement field $\vec{S}$ defined by \ref{mdof_displacement} and the closed counting form $\omega$ defined by \ref{mdof_form} such that $\omega$ is killed only by the displacement field $\vec{S}$ as in \ref{mdof_displacement_kills} or by a field everywhere tangent to $\vec{S}$.
-	\item show that the potential $\theta$ for $\omega$ defined in \ref{mdof_form_potential} can be expressed, without loss of generality, as \ref{mdof_potential_expression} and that $\omega$ can be expressed as $\ref{mdof_form_components}$
-	\item show that \ref{mdof_form_potential}, \ref{mdof_displacement_kills} and \ref{mdof_potential_expression} recover Hamilton's equations as in \ref{mdof_Ham_eq}
-	\item show that the integral of $\omega$ through a surface $\Sigma$ delimited by a path $\gamma$ and its variation $\gamma'$ is the variation of the integral of the vector potential, as claimed in \ref{mdof_action}
-	\item show that the integral of the vector potential is stationary for and only for the solutions of the equations of motion
-	\item show that the integral of the vector potential reduces to the action if and only if \ref{assum_kineq} applies.
-\end{enumerate}
-
-For (i), the arguments presented for the single d.o.f. still apply: \ref{assum_detrev} leads to the existence of a displacement field $\vec{S}$
-\begin{equation}
-\tag*{(\ref{mdof_displacement})}
-	\begin{aligned}
-		\vec{S} &= S^a e_a = \frac{d\xi^a}{dt} e_a =\frac{dq^i}{dt} e_{q^i} + \frac{dp_i}{dt} e_{p_i} + \frac{dt}{dt} e_t
-	\end{aligned}
-\end{equation}
-that tells us how states move in the extended phase space.
-We also need to reformulate state/density conservation in time, which we cannot do simply with a surface integral of $\vec{S}$.
-To this end, we need a functional $\mathcal{F}$ that takes a hyper-surface $\Sigma$ in the extended phase space and returns the count of states.
-We have
-\begin{equation}
-	\mathcal{F}(\Sigma) = \int_\Sigma \Omega(d\Sigma)
-\end{equation}
-where $\Omega$ is the related form.
-Given the one-to-one correspondence between states and evolutions under \ref{assum_detrev}, $\Omega$ can be understood as quantifying either the flow of states through the surface or the states that are present on the surface.
-
-As in the previous case, assumption \ref{assum_detrev} tells us that the integral over a closed hyper-surface is zero, since as many states flow in as flow out. We have
-\begin{equation}
-	\begin{aligned}
-		\oint_{\Sigma = \partial V} \Omega(d\Sigma) &= 0 = \int_V \partial \Omega(dV) \\
-		\partial\Omega &= 0.
-	\end{aligned}
-\end{equation}
-which tells us that $\Omega$ is closed.
-
-Given that $\Omega$ quantifies the flow of $\vec{S}$ through $\Sigma$, the flow for any surface that is tangent to $\vec{S}$ must vanish.
-If $d\gamma = \vec{S} \varphi dt$ for some scalar $\varphi$, we must have
-\begin{equation}
-	\Omega(d\gamma, \cdot) = \varphi dt \Omega(\vec{S}, \cdot) = 0.
-\end{equation}
-The notation is a bit improper as $\vec{S}$ is not technically an infinitesimal displacement, but a rate of displacement given a parameter.
-However, the slight abuse of notation makes the concept more clear: the direction of motion does not contribute to the flow through the surface $\Sigma$.
-
-On the other hand, this must be the only direction that does not contribute to the flow.
-Since $\vec{S}$ cannot be zero because states must flow in time, for any other direction $\vec{T}$ that is not parallel to $\vec{S}$ there must be a surface tangent to $\vec{T}$ and not $\vec{S}$.
-The flow through that surface cannot be zero.
-Therefore the only way that the flow is zero for any surface tangent to a vector $\vec{T}$ is if $\vec{T}= \varphi \vec{S}$
-
-We now turn to assumption \ref{assum_indep}.
-This requires the count of states to factorize: if we have a parallelepiped in phase space, the count of states (i.e.
-its volume) is the product of the count of the configurations over each independent d.o.f.
-(i.e.
-the area of the sides).
-Equivalently, \ref{assum_indep} must allow the case of a distribution over states with no correlations between the independent d.o.f., in which case the density must factorize.
-Since we must be able to quantify the configuration for each d.o.f., and since each d.o.f.
-is two dimensional, we must have a two-form $\omega$ such that:
-\begin{equation}
-	\Omega = \omega^N.
-\end{equation}
-
-Using assumptions \ref{assum_detrev} and \ref{assum_indep} again we can see how $\omega$ must be closed
-\begin{equation}
-	\partial \omega = 0.
-\end{equation}
-If we imagine translating a surface along an independent variable, the number of configurations identified by the surface cannot change.
-This uses \ref{assum_indep}.
-If we imagine translating in time, again the number of configurations cannot change because of \ref{assum_detrev}.
-Therefore, if we imagine a parallelepiped, opposite sides must identify the same number of configurations.
-The integral will take the sides with an opposite orientation, and therefore there will be zero net contribution.
-It is interesting to note that this condition requires both assumptions.
-
-Putting it all together, assuming \ref{assum_detrev} and \ref{assum_indep} gives us the existence of the displacement vector field $\vec{S}$ and the closed counting two-form $\omega$ such that $\vec{S}$ identifies the only vector field that kills $\omega$.
-
-The converse is also true: if we start with the mathematical requirements, we can recover the physical assumptions.
-If we are given a two-form $\omega$ on a $2N+1$ dimensional space, we can define the $2N$ form $\Omega = \omega^N$.
-This form will be non-zero because $\vec{S}$ is the only vector field that kills the two-form $\omega$.
-It will also be closed since $\partial \omega = 0$ and
-\begin{equation}
-	\partial \Omega = N\partial \omega \wedge \omega^{N-1} = 0.
-\end{equation}
-So the flow of the well defined displacement field $\vec{S}$, as measured by $\Omega$, is zero through closed surfaces. This implies \ref{assum_detrev}.
-
-To obtain \ref{assum_indep}, if each pair of state variables $q^i$, $p_i$ forms an independent degree of freedom, only the surfaces spanned by matching position and momentum will define actual configurations, while all other combinations will not define any.
-Therefore:
-\begin{equation}\label{canonical_conditions}
-	\begin{aligned}
-		\omega(e^{q^i}, e^{p_j}) = \omega_{q^i p_j} &= \delta^i_j = - \omega_{p_j q^i} \\
-		\omega(e^{q^i}, e^{q^j}) = \omega_{q^i q^j} &= 0 \\
-		\omega(e^{p_i}, e^{p_j}) = \omega_{p_i p_j} &= 0
-	\end{aligned}
-\end{equation}
-If we start with a closed form $\omega$ that has a single degenerate direction, by Darboux's theorem we can find coordinates for the non-degenerate part of $\omega$ that satisfy the above conditions.
-This recovers \ref{assum_indep}.
-
-Note that we have broken down exactly how each assumption contributes to the mathematical setting.
-If \ref{assum_indep} fails, then $\Omega$ cannot be understood as $\omega^N$, which means we have much fewer conditions on the motion: we only know that the total flow is conserved, but we cannot break it down along each d.o.f..
-On the other hand, if \ref{assum_detrev} fails we may not have a single displacement field $\vec{S}$ and/or the form that counts states would not be closed.
-Nonholonomic constraints will, in general, make one of the assumptions fail, for example by introducing a dependence between degrees of freedom.
-This type of exact mapping between assumptions and their mathematical expression is one of the key goals of Reverse Physics in particular and our overall project Assumptions of Physics more in general.
-
-For (ii), we note that since the form is closed, we can find, at least in a contractible region, a covector potential $\theta$ such that $\omega=\partial \theta$.
-Those familiar with symplectic and contact geometry will know that using Darboux's theorem, we can find a set of coordinates such that we can express $\theta$ as
-\begin{equation}
-	\tag*{(\ref{mdof_potential_expression})}
-	\theta = p_i e^{q^i} - H e^t.
-\end{equation}
-Just stating this is, of course, not very insightful.
-To gain more insight, we provide the sketch of a proof that, as we have done for the single degree of freedom, uses gauge transformations to fix the form of the potential.
-
-As a starting point, we take conditions \ref{canonical_conditions} that state mathematically that $q^i$ and $p_i$ form independent degrees of freedom.
-We re-express those conditions in terms of the potential $\theta$
-\begin{equation}\label{canonical_potential_conditions}
-\begin{aligned}
-	\omega(e^{q^i}, e^{p_j}) &= (-\partial\theta)_{q^i p_j} = -(\partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i}) = \delta^i_j \\
-	\omega(e^{q^i}, e^{q^j}) &= (-\partial\theta)_{q^i q^j} = -(\partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i}) = 0 \\
-	\omega(e^{p_i}, e^{p_j}) &= (-\partial\theta)_{p_i p_j} = -(\partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i}) = 0
-\end{aligned}
-\end{equation}
-We can use our gauge freedom to set $\theta_{p_1} = 0$, much in the same way we did for the simpler case.
-We now have $\partial_{q^1} \theta_{p_1} = 0$ and, by the first condition, $\partial_{p_1} \theta_{q^1} = 1$.
-Integrating, we have $\theta_{q^1} = p_1 + g(q^i, p_2, p_3, ..., t)$ where $g$ is an arbitrary function which we can set to zero.
-Like in the single degree of freedom, we can do this because $g$ does not depend on $p_1$, so it can be eliminated with a gauge transformation that does not change $\theta_{p_1}$.
-Therefore we have:
-\begin{equation}
-	\theta = p_1 e^{q^1} + 0 e^{p_1} + \theta_{q^2} e^{q^2} + \theta_{p_2} e^{p_2} + ... + \theta_{t} e^{t}.
-\end{equation}
-
-Note that the components for the first degree of freedom do not depend on the other degrees of freedom.
-That is, for all $i>1$, $\partial_{q^i} \theta_{q^1} = \partial_{p_i} \theta_{q^1} = \partial_{q^i} \theta_{p_1} = \partial_{p_i} \theta_{p_1} = 0$.
-But by using conditions \ref{canonical_potential_conditions}, we find that the converse is true as well: the components of all other degrees of freedom do not depend on the first.
-That is, for all $i>1$, $\partial_{q^1} \theta_{q^i} = \partial_{p_1} \theta_{q^i} = \partial_{q^1} \theta_{p_i} = \partial_{p_1} \theta_{p_i} = 0$.
-
-We can then use, again, our gauge freedom with a function that does not depend on the first two variables to set $\theta_{p_2} = 0$.
-And, with the same reasoning, we will be able to set $\theta_{q^2} = p_2$.
-And then, again, find that the first two degrees of freedom do not depend on the others, etc.
-At the end, we will find \ref{mdof_potential_expression}.
-
-The insight is that the independence of the degrees of freedom leads to the independent expression of the components of the covector potential.
-We stress again that this is a choice, and that the specific expression of the components is not physically constrained.
-
-The components of $\partial\theta$ are
-\begin{equation}
-	\begin{aligned}
-		(\partial\theta)_{q^i p_j} &= \partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i} = \partial_{q^i} 0 - \partial_{p_j} p_i= - \delta^i_j \\
-		(\partial\theta)_{q^i q^j} &= \partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i} = \partial_{q^i}p_j - \partial_{q^j}p_i = 0 \\
-		(\partial\theta)_{p_i p_j} &= \partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i} = \partial_{p_i} 0 - \partial_{p_j} 0 = 0 \\
-		(\partial\theta)_{q^i t} &= \partial_{q^i}\theta_{t} - \partial_{t}\theta_{q^i} = \partial_{q^i} (-H) - \partial_{t} p_i= - \partial_{q^i} H \\
-		(\partial\theta)_{p_i t} &= \partial_{p_i}\theta_{t} - \partial_{t}\theta_{p_i} = \partial_{p_i} (-H) - \partial_{t} 0= - \partial_{p_i} H \\
-	\end{aligned}
-\end{equation}
-Therefore we recover
-\begin{equation}
-	\tag*{(\ref{mdof_form_components})}
-	\omega_{ab}=(-\partial\theta)_{ab} = \begin{bmatrix}
-		0 & \delta^j_i & \partial_{q^i} H \\
-		-\delta^i_j & 0 & \partial_{p_i} H \\
-		- \partial_{q^j} H & - \partial_{p_j} H & 0
-	\end{bmatrix}.
-\end{equation}
-
-For (iii) we calculate $\omega(\vec{S}, \cdot ) $. We have
-\begin{equation}
-\begin{aligned}
-	\omega(\vec{S}, \cdot )  &= S^a \omega_{ab} e^b = 0 \\
-	&= (S^{q^i}\omega_{q^ib} + S^{p_i}\omega_{p_ib} + S^{t}\omega_{tb}) e^b \\
-	&= (S^{q^i}\omega_{q^iq^j} + S^{p_i}\omega_{p_iq^j} + S^{t}\omega_{tq^j}) e^{q^j} + \\
-	& (S^{q^i}\omega_{q^ip_j} +  S^{p_i}\omega_{p_ip_j} + S^{t}\omega_{tp_j}) e^{p_j} + \\
-	& (S^{q^i}\omega_{q^it} + S^{p_i}\omega_{p_it} + S^{t}\omega_{tt}) e^t \\
-	&= (-S^{p_i}\delta^i_j - S^{t}\partial_{q^j} H ) e^{q^j} + \\
-	& (S^{q^i}\delta^j_i -  S^{t}\partial_{p_j} H) e^{p_j} + \\
-	& (S^{q^i} \partial_{q^i} H + S^{p_i} \partial_{p_i} H) e^t \\
-\end{aligned}
-\end{equation}
-
-All components must be zero, therefore we have the following three equations:
-\begin{equation}
-\begin{aligned}
-	S^{p_j} &= - S^{t} \partial_{q^j} H \\
-	S^{q^j} &= S^{t}\partial_{p_j} H \\
-	S^{q^i} \partial_{q^i} H + S^{p_i} \partial_{p_i} H &= S^{t}\partial_{p_i} H \partial_{q^i} H \\
-	& - S^{t} \partial_{q^i} H \partial_{p_i} H = 0
-\end{aligned}
-\end{equation}
-Note that the last expression is not a new equation: it is identical to zero given the previous two equations. Since $S^t = 1$, we have
-\begin{equation}
-	\tag*{(\ref{mdof_Ham_eq})}
-\begin{aligned}
-	S^{q^j} &= \partial_{p_j} H \\
-	S^{p_j} &= - \partial_{q^j} H 
-\end{aligned}
-\end{equation}
-which recovers Hamilton's equations for multiple degrees of freedom.
-
-The argument can be run in the other direction.
-If we have a system that obeys \ref{mdof_Ham_eq}, we can define the two-form $\omega$ using \ref{mdof_form_components}.
-We can verify that $\omega$ is closed, it is killed by $\vec{S}$ and recover all the starting points.
-Therefore we find again complete equivalence with Hamiltonian mechanics and assumptions \ref{assum_detrev} and \ref{assum_indep}.
-
-For those familiar with general relativity, recall that the metric tensor $g_{\alpha\beta}$ defines the geometry by defining lengths and angles.
-Here we find something similar in the sense that the two-form $\omega_{ab}$ is a rank two tensor that defines the geometry of the space.
-The difference is that $\omega_{ab}$ does not define lengths and angles, but only areas of two-dimensional surfaces, which corresponds to the number of configurations on each independent degree of freedom.
-Mathematically, this difference is between the symmetry of $g_{\alpha\beta} = g_{\beta\alpha}$ and the anti-symmetry of $\omega_{ab} = -\omega_{ba}$.
-In this context, what the equations of motion say is that states move in time at the same rate (i.e.
-$\frac{dt}{dt} = 1$) with a deterministic and reversible flow (i.e.
-no states are created or destroyed on any independent degree of freedom).
-Nothing else.
-
-Comparing \ref{mdof_form_components} and \ref{mdof_Ham_eq} we see that the time components $\omega_{at}$ are actually the components of $\vec{S}$.
-In fact, the only non-trivial components of $\omega$ are components of $\vec{S}$.
-These two objects, then, are not simply related: they express the same information.
-
-For a single degree of freedom, recalling that $S^t = 1$, we can describe $\omega$ with the following nine components
-\begin{equation}\label{mdof_reduceto_sdof}
-	\omega_{ab}=(-\partial\theta)_{ab} = \begin{bmatrix}
-		0 & S^t & -S^p \\
-		-S^t & 0 & S^q \\
-		S^p & - S^q & 0
-	\end{bmatrix}.
-\end{equation}
-In three dimensions, using the Levi-Civita symbol $\epsilon^{abc}$ we can write
-\begin{equation}
-	S^a = \epsilon^{abc} \omega_{bc}
-\end{equation}
-which tells us that $\vec{S}$ is the pseudo-vector corresponding to the direction perpendicular to $\omega$.
-This is the same relationship that we would have, for example, for the magnetic field psuedo-vector and the spatial components of the electromagnetic field in relativity.
-If we express $\omega$ in terms of $\vec{\theta}$, we have
-\begin{equation}
-	S^a = - \epsilon^{abc} \partial_b \theta_c
-\end{equation}
-which recovers \ref{sdof_displacement_potential}: the displacement field is minus the curl of $\omega$. Thus we can see exactly how the general case reduces to the single d.o.f. case.
-
-For (iv), we simply need to apply Stokes' theorem.
-The situation is identical to the single degree of freedom, except that the flow now is not given by the surface integral of $\vec{S}$, but of the form $\omega$.
-The path $\gamma$ and its variation $\gamma'$ will form a closed curve where $\gamma'$ is taken with the opposite orientation.
-Because $\omega$ is closed, its surface integral depends only on its contour.
-Moreover, by Stokes' theorem, the flow through the surface will be equal to the integral of the potential over the contour.
-Putting it all together we recover
-\begin{equation}
-\tag*{(\ref{mdof_action})}
-\begin{aligned}
-	- \int_{\Sigma} \omega(d\Sigma) &=
-	\int_{\Sigma} \partial \theta (d\Sigma) \\
-	&= \oint_{\partial \Sigma = \gamma \cup \gamma'} \theta(d\gamma) \\
-	&= \int_{\gamma} \theta(d\gamma) - \int_{\gamma'} \theta (d\gamma') \\
-	&= \delta \int_{\gamma} \theta (d\gamma).
-\end{aligned}
-\end{equation}
-
-For (v), if the path $\gamma$ is a solution to the equations of motion, it is tangent to $\vec{S}$ at every point.
-Therefore we can write $d\gamma= \vec{S} dt$.
-Each infinitesimal surface element $d\Sigma$ will be an infinitesimal parallelogram with $d\gamma$ as one side, and the other side $d\lambda$ connects the path to the variation.
-We have:
-\begin{equation}
-\tag*{(\ref{mdof_stationary_action})}
-\begin{aligned}
-	\delta \int_{\gamma} \theta (d\gamma) &= -\int_\Sigma \omega(d\Sigma) \\
-	&= -\int_\Sigma \omega(d\gamma, d\lambda) \\
-	&= -\int_\Sigma \omega(\vec{S}, d\lambda) dt = 0.
-\end{aligned}
-\end{equation}
-The converse is also true: suppose that, for a fixed $\gamma$, the flow through $\Sigma$ is zero no matter the choice of $\gamma'$.
-Then at every point $\omega(d\gamma, \cdot) dt = 0$, which means $d\gamma$ kills $\omega$.
-But since $\vec{S}$ is the only direction that kills $\omega$ at every point, $d\gamma$ must be everywhere tangent to $\vec{S}$ and therefore $\gamma$ is a solution of the equations of motion.
-
-For (vi), the argument is exactly the same as in the single d.o.f..
-As the evolution is differentiable, the velocity $\dot{q}^i = \frac{dq^i}{dt}$ is well defined.
-Under \ref{assum_kineq} there exists an invertible map $p_i=p_i(q^j, \dot{q}^j, t)$ that allows us to reconstruct the momentum at a given time by knowing position and velocity.
-We can therefore express the principle of stationary action only in terms of position and velocity.
-We have
-\begin{equation}
-\tag*{(\ref{mdof_Lagrangian})}
-	\begin{aligned}
-		\delta \int_\gamma \theta (d\gamma) &= \delta \int^{t_2}_{t_1} \theta \left(\frac{d\gamma}{dt} dt\right) \\
-		&= \delta \int^{t_2}_{t_1} \left(\theta_{q^i} \frac{dq^i}{dt} + \theta_{p_i} \frac{dp_i}{dt} + \theta_t \frac{dt}{dt}\right) dt \\
-		&= \delta\int^{t_2}_{t_1} \left(p_i \frac{dq^i}{dt} + 0 \frac{dp_i}{dt} - H \frac{dt}{dt}\right) dt \\
-		&= \delta \int^{t_2}_{t_1} \left(p_i \frac{dq^i}{dt} - H\right) dt \\
-		&= \delta \int^{t_2}_{t_1} L(q^i, \dot{q}^i, t) \, dt = 0.
-	\end{aligned}
-\end{equation}
-This recovers the principle of stationary action in standard form.
-
-Again, the argument can proceed in the opposite direction.
-If we have a system for which the principle of stationary action yields a unique solution, we can write $p_i = \frac{\partial L}{\partial \dot{q}^i} = p_i(q^j, \dot{q}^j, t)$.
-This means that we can reconstruct the momentum from the kinematics, and therefore \ref{assum_kineq} is satisfied.
-Moreover, we can define a Hamiltonian $H = p_i \dot{q}^i - L$ such that the paths that satisfy the equations of motion \ref{mdof_Ham_eq} are exactly those that satisfy the principle of stationary action.
-Therefore the system satisfies \ref{assum_detrev} and \ref{assum_indep} as well.
-
-The comments on the lack of strict physicality for the values of the Lagrangian and the action remain unchanged.
-They both depend on a choice of gauge.
-The comment on the centrality of the extended phase space to understand the action principle is only reinforced.
-Additionally, note that by changing $t$ to $ct$ expression \ref{mdof_potential_expression} becomes
-\begin{equation}\label{mdof_potential_relativistic}
-	\theta = p_i e^{q^i} - \frac{H}{c} e^{ct},
-\end{equation}
-which bears striking resemblance to the relativistic four-momentum.
-While it is outside the scope of this article, the idea is that, in this setting, we already have some pre-relativistic features, even though we have not introduced a metric tensor.
-
-We hope that this result shows that it is possible to bring physical ideas and mathematical concepts much closer together in a way that is more clear and satisfying.
-
 \end{document}

### 2023-06-27 Articles/2022-ActionPrinciple/ScientificReportsAmended/ActionPrincipleSR.tex
ActionPrinciple: Scientific Reports version

--- 2022-10-13 Articles/2022-ActionPrinciple/ActionPrincipleArchive.tex
@@
 
-\section{Introduction}
+\section*{Introduction}
 
@@
 We then proceed to the case of multiple independent degrees of freedom using tools from differential geometry.
-The main article will presents all the key point needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the appendix.
-
-\section{One degree of freedom}
+The main article will present all the key point needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the Supplementary Information.
+
+\section*{One degree of freedom}
 
@@
 \end{equation}
-that describes how states move in time.\footnote{Where possible, we will be writing the same expression in both vector calculus and component notations.} The time component of the displacement vector field is constrained, as we have
+that describes how states move in time.
+[Where possible, we will be writing the same expression in both vector calculus and component notations.] The time component of the displacement vector field is constrained, as we have
 \begin{equation}\label{sdof_time_constraint}
@@
 
-Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative\footnote{We write the exterior derivative as $\partial \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.} of a covector $\theta$:
+Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative of a covector $\theta$:
 \begin{equation}\label{mdof_form_potential}
@@
 \end{equation}
-The minus sign is required to match conventions.
+The minus sign is required to match conventions. [We write the exterior derivative as $\partial \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.]
 
@@
 
+\vspace{-1mm}
+\section*{Data availability}
+Not applicable.
+
 \bibliography{bibliography}
 
-\section*{Appendix}
-
-In this section we will review the arguments in more detail.
+\newpage
+\section*{Supplementary Information}
+\setcounter{page}{1}
+
+In this Supplementary Information we will review the arguments in more detail.
 Given that this article is still aimed at a physics audience, the focus will be conceptual rigor, rather than mathematical rigor.
@@
 
-We divide the appendix into three sections: the first will address the single d.o.f.
-case, the second will give a geometric interpretation of the tools of differential geometry that is suitable for physicists, and lastly we address the multiple d.o.f.
+We divide the material into three sections: the first will address the single d.o.f.
+case, the second will give a geometric interpretation of the tools of differential topology that is suitable for physicists, and lastly we address the multiple d.o.f.
 case.
@@
 \end{equation}
+In the literature of dynamical systems, the displacement vector field $\vec{S}$ is simply called the vector field of the dynamical system.
+The expression of the vector field is what determines the equations of motion in a dynamical system.
 The double role of time gives us
@@
 
+The potential $\vec{\theta}$ is closely related to the canonical one-form of symplectic geometry and the contact form of contact geometry, but there are some differences.
+Both the canonical one-form and the contact form are a specific covector, defined with specific coordinates.
+For us, $\vec{\theta}$ is a potential and therefore is not fully specified by the physics.
+Writing down a specific potential means choosing a gauge, therefore our potential corresponds to the canonical one-form or the contact form only if a specific gauge is fixed.
+
 For (iii), we expand each component of \ref{sdof_displacement_potential} using \ref{sdof_potential_expression}.
@@
 Moreover, by Stokes' theorem, the flow through the surface will be equal to the integral of the potential over the contour.
-Putting it all together we have
+Thus we have
 \begin{equation}
@@
 
-Now, since every functional has a corresponding form, $\partial F$ must have a corresponding $k+1$-form which must be fully identified by $\omega$. We note it as $\partial \omega$ and write
-\begin{equation*}
-	\partial F(\Sigma) = \int_\Sigma \partial \omega(d\Sigma) = \oint_{\sigma = \partial \Sigma} \omega(d\sigma) = F(\partial \Sigma).
-\end{equation*}
-This is the generalized Stokes' theorem and $\partial \omega$ is the exterior derivative of $\omega$. We use $\partial$ as we leave $d$ reserved for the infinitesimal regions.
+Now, since every functional has a corresponding form, $\partial F$ must have a corresponding $k+1$-form which must be fully identified by $\omega$. We note it as $\partial \wedge \omega$ and write
+\begin{equation*}
+	\partial F(\Sigma) = \int_\Sigma \partial \wedge \omega(d\Sigma) = \oint_{\sigma = \partial \Sigma} \omega(d\sigma) = F(\partial \Sigma).
+\end{equation*}
+This is the generalized Stokes' theorem and $\partial \wedge \omega$ is the exterior derivative of $\omega$. We use $\partial$ as we leave $d$ reserved for the infinitesimal regions.
 
@@
 \begin{equation*}
-	\partial \partial \omega = 0,
+	\partial \wedge \partial \wedge \omega = 0,
 \end{equation*}
@@
 
-In the same way that a divergence-free field can be written as the curl of a vector potential, or a curl-free field can be written as the gradient of a scalar potential, a form whose exterior derivative is zero $\partial \omega = 0$ (a closed form), can be written as the exterior derivative $\omega = \partial \theta$ of a lower dimensional form, at least in some neighborhood.
+In the same way that a divergence-free field can be written as the curl of a vector potential, or a curl-free field can be written as the gradient of a scalar potential, a form whose exterior derivative is zero $\partial \wedge \omega = 0$ (a closed form), can be written as the exterior derivative $\omega = \partial \wedge \theta$ of a lower dimensional form, at least in some neighborhood.
 In terms of functionals, this corresponds to the case where $F(\sigma)=0$ for all contractible surfaces, those surfaces that can be morphed to a point with a continuous transformation.
@@
 	\begin{aligned}
-		\oint_{\Sigma = \partial V} \Omega(d\Sigma) &= 0 = \int_V \partial \Omega(dV) \\
-		\partial\Omega &= 0.
+		\oint_{\Sigma = \partial V} \Omega(d\Sigma) &= 0 = \int_V \partial \wedge \Omega(dV) \\
+		\partial \wedge\Omega &= 0.
 	\end{aligned}
@@
 \begin{equation}
-	\partial \omega = 0.
+	\partial \wedge \omega = 0.
 \end{equation}
@@
 
-Putting it all together, assuming \ref{assum_detrev} and \ref{assum_indep} gives us the existence of the displacement vector field $\vec{S}$ and the closed counting two-form $\omega$ such that $\vec{S}$ identifies the only vector field that kills $\omega$.
+The form $\omega$ corresponds to the symplectic form of symplectic manifolds and of the symplectic bundles of contact manifolds with the following caveats.
+As we are in phase space extended by time, we have a contact manifold, and $\omega$ mathematically corresponds to the exterior derivative of the contact form, which is the primitive mathematical object.
+Physically, this is backwards: $\omega$ is the actual primitive physical object and the contact form $\theta$ is a potential, which is unphysical as it is fixed by an arbitrary gauge.
+Note how the mathematical definitions make it hard to have a coherent physical interpretation.
+The primitive objects for contact manifolds and symplectic manifolds are different, $\theta$ in the first case and $\omega$ in the other, and they are defined and studied as two separate mathematical spaces, while physically they are the same space with or without time.
+
+To summarize, assuming \ref{assum_detrev} and \ref{assum_indep} gives us the existence of the displacement vector field $\vec{S}$ and the closed counting two-form $\omega$ such that $\vec{S}$ identifies the only vector field that kills $\omega$.
 
@@
 This form will be non-zero because $\vec{S}$ is the only vector field that kills the two-form $\omega$.
-It will also be closed since $\partial \omega = 0$ and
-\begin{equation}
-	\partial \Omega = N\partial \omega \wedge \omega^{N-1} = 0.
+It will also be closed since $\partial \wedge \omega = 0$ and
+\begin{equation}
+	\partial \wedge \Omega = N\partial \wedge \omega \wedge \omega^{N-1} = 0.
 \end{equation}
@@
 
-For (ii), we note that since the form is closed, we can find, at least in a contractible region, a covector potential $\theta$ such that $\omega=\partial \theta$.
+For (ii), we note that since the form is closed, we can find, at least in a contractible region, a covector potential $\theta$ such that $\omega=\partial \wedge \theta$.
 Those familiar with symplectic and contact geometry will know that using Darboux's theorem, we can find a set of coordinates such that we can express $\theta$ as
@@
 \begin{aligned}
-	\omega(e^{q^i}, e^{p_j}) &= (-\partial\theta)_{q^i p_j} = -(\partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i}) = \delta^i_j \\
-	\omega(e^{q^i}, e^{q^j}) &= (-\partial\theta)_{q^i q^j} = -(\partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i}) = 0 \\
-	\omega(e^{p_i}, e^{p_j}) &= (-\partial\theta)_{p_i p_j} = -(\partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i}) = 0
+	\omega(e^{q^i}, e^{p_j}) &= (-\partial \wedge\theta)_{q^i p_j} = -(\partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i}) = \delta^i_j \\
+	\omega(e^{q^i}, e^{q^j}) &= (-\partial \wedge\theta)_{q^i q^j} = -(\partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i}) = 0 \\
+	\omega(e^{p_i}, e^{p_j}) &= (-\partial \wedge\theta)_{p_i p_j} = -(\partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i}) = 0
 \end{aligned}
@@
 
-The components of $\partial\theta$ are
-\begin{equation}
-	\begin{aligned}
-		(\partial\theta)_{q^i p_j} &= \partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i} = \partial_{q^i} 0 - \partial_{p_j} p_i= - \delta^i_j \\
-		(\partial\theta)_{q^i q^j} &= \partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i} = \partial_{q^i}p_j - \partial_{q^j}p_i = 0 \\
-		(\partial\theta)_{p_i p_j} &= \partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i} = \partial_{p_i} 0 - \partial_{p_j} 0 = 0 \\
-		(\partial\theta)_{q^i t} &= \partial_{q^i}\theta_{t} - \partial_{t}\theta_{q^i} = \partial_{q^i} (-H) - \partial_{t} p_i= - \partial_{q^i} H \\
-		(\partial\theta)_{p_i t} &= \partial_{p_i}\theta_{t} - \partial_{t}\theta_{p_i} = \partial_{p_i} (-H) - \partial_{t} 0= - \partial_{p_i} H \\
+The components of $\partial \wedge\theta$ are
+\begin{equation}
+	\begin{aligned}
+		(\partial \wedge\theta)_{q^i p_j} &= \partial_{q^i}\theta_{p_j} - \partial_{p_j}\theta_{q^i} = \partial_{q^i} 0 - \partial_{p_j} p_i= - \delta^i_j \\
+		(\partial \wedge\theta)_{q^i q^j} &= \partial_{q^i}\theta_{q^j} - \partial_{q^j}\theta_{q^i} = \partial_{q^i}p_j - \partial_{q^j}p_i = 0 \\
+		(\partial \wedge\theta)_{p_i p_j} &= \partial_{p_i}\theta_{p_j} - \partial_{p_j}\theta_{p_i} = \partial_{p_i} 0 - \partial_{p_j} 0 = 0 \\
+		(\partial \wedge\theta)_{q^i t} &= \partial_{q^i}\theta_{t} - \partial_{t}\theta_{q^i} = \partial_{q^i} (-H) - \partial_{t} p_i= - \partial_{q^i} H \\
+		(\partial \wedge\theta)_{p_i t} &= \partial_{p_i}\theta_{t} - \partial_{t}\theta_{p_i} = \partial_{p_i} (-H) - \partial_{t} 0= - \partial_{p_i} H \\
 	\end{aligned}
@@
 	\tag*{(\ref{mdof_form_components})}
-	\omega_{ab}=(-\partial\theta)_{ab} = \begin{bmatrix}
+	\omega_{ab}=(-\partial \wedge\theta)_{ab} = \begin{bmatrix}
 		0 & \delta^j_i & \partial_{q^i} H \\
@@
 \begin{equation}\label{mdof_reduceto_sdof}
-	\omega_{ab}=(-\partial\theta)_{ab} = \begin{bmatrix}
+	\omega_{ab}=(-\partial \wedge\theta)_{ab} = \begin{bmatrix}
 		0 & S^t & -S^p \\
@@
 Moreover, by Stokes' theorem, the flow through the surface will be equal to the integral of the potential over the contour.
-Putting it all together we recover
+Thus we have
 \begin{equation}
@@
 	- \int_{\Sigma} \omega(d\Sigma) &=
-	\int_{\Sigma} \partial \theta (d\Sigma) \\
+	\int_{\Sigma} \partial \wedge \theta (d\Sigma) \\
 	&= \oint_{\partial \Sigma = \gamma \cup \gamma'} \theta(d\gamma) \\

### 2023-06-27 Articles/2022-ActionPrinciple/ScientificReportsAmended/response1.tex
ActionPrinciple: Scientific Reports version

\documentclass[11pt, executivepaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={158mm,237mm},
 left=28mm,
 top=28mm,
 }
\usepackage{hyperref}
\hypersetup{colorlinks= true, allcolors=blue}
\setcitestyle{aysep={}}
\begin{document}

\title{\textbf{Replies to Reviewers' Comments}}

\author{The Authors}

\maketitle

\noindent We would like to thank the reviewers for their time and comments.

\section{Replies to Reviewer 1}

\begin{quote}
 In this paper the authors are studying the Principle of Least Action (PLA) from a geometrical point of view and trying to provide a physical meaning for the principle. The paper relies on analytical mechanics definition and differential geometry tools.

The authors introduced the current effort as part of a main project. The main project is named Reverse Physics, at which a set of starting physical assumptions that are sufficient to rederive current theories are derived. They considered Lagrangian mechanics (i.e., PLA) as a candidate for this study. They claim that they found that the PLA is equivalent to three assumptions: 1) determinism/reversibility, 2) independence of degrees of freedom and 3) kinematics/dynamics equivalence for any divergence-free fields. The authors claims that the mathematics for understanding the PLA is well established in literature, however, a clear geometric interpretation of the PLA along with a tight connection between the math and physics are lacking, hence, providing these two lacking points constructs the premises of this paper.

They considered a dynamical system with one degree of freedom (DOF), to simplify the analysis and the presentation of the idea. Then they switched to systems with higher DOF, at which the differential geometry tools are needed to proceed further. The study relies on the generalized phase-space constructed from the generalized co-ordinates and momenta (p,q,t). Following the three assumptions, the evolution of any dynamical system in the phase-space could be considered as a divergence free flow problem (i.e., incompressible fluid problem).
\end{quote}
We are glad that the referee understood the premise of the paper.

\begin{quote}
 The paper is interesting to the community and provides an intriguing geometrical representation of an analytical principle.  However, there are several concerns (listed below) that need to be addressed before publication.

 The authors claim to provide “a clear geometric interpretation of the action principle”. However, this geometric interpretation is inherently present in the analytical mechanics field. For example, when they mentioned in pp.3 “Now suppose $\gamma$ is a solution of the … paths for which”, they described a geodesic (i.e., shortest distance over curved manifolds) between two points on the manifold of phase-space, which is an already established fact.
\end{quote}
It is not clear to us to what geometric interpretation the reviewer is referring. Geodesics as shortest distance can only be defined on spaces where distance is defined, which are Riemannian manifolds in the context of differential geometry. The spaces we are considering are either symplectic (e.g. phase space) or contact manifolds (e.g. phase space extended by time). There is no notion of distance on these spaces, therefore no geodesic can be defined. We looked in the literature for possible analogs of geodesics; we found none for contact manifolds and we found pseudoholonomic curves for symplectic manifolds, though these do not appear to have relevance.

\begin{quote}
 The authors claim to provide “a tight connection between the math and the physics the action principle represents”. But there is a counter argument for this statement. The PLA, Newtonian mechanics, Gauss principle ... etc., all of these are principles or could be thought of as axioms. They do not have a physical meaning by themselves; however, we (researchers) use them to understand the physics of nature. The authors introduced three assumptions (DR, KE \& IND) and claimed that these assumptions explained the PLA physically. However, in reality, these assumptions forced the phase-space to be a divergence free flow filed, hence, helped in formulating and understanding the geometrical representation of the PLA. Yet, after the study, the main logic or philosophy behind the PLA remains to be the minimization of the action integral (i.e., minimization of a certain measure/gauge quantity). See the below table for more details.

 Newtonian mechanics \\
 $\mathit{F}=m\mathit{a}$ \\
 The axioms are Newton’s three laws.

 Lagrangian mechanics \\
 $\delta\int_{t_1}^{t_2}L\left(q,\dot{q},t\right)dt=0$ \\
 The axiom for the PLA is dated back to Maupertuis principle. Where nature minimizes a magical quantity (gauge).

 Gauss Principle ~ Principle of Least Curvature \\ 
 $Z=\sum_{j=1}^{n}m_j\cdot\left|\ddot{r_j}-\frac{F_j}{m_j}\right|^2$ \\
 The axiom is that mechanical systems must evolve with time such the constraint force (i.e., accelerations) are minimum.

 All these principles do not have physics, we use them to study physics. Nevertheless, this does not mean, one can be presented geometrically the other cannot. 
\end{quote}
It is not clear to us why one would say that Newton’s three laws do not have physical meaning by themselves. They are all in terms of operationally defined, measurable and physically intuitive objects. The second law is particularly physically meaningful as it stands in contrast to Aristotelian thinking that posited that a constant force would give constant velocity. In this regard, this is exactly why the principle of stationary action is lacking physical motivation, as it is expressed in terms of a quantity (the action) that is not operationally defined, measurable or physically intuitive. The reviewer is correct in stating that, after our study, the main logic behind the principle is the same. But we have clarified exactly that the action is \textbf{not} a physical quantity, as it is gauge dependent, and why the variation of the action \textbf{is} a physical quantity, and how it relates to the flow (i.e. surface integral) of the displacement field, which is a physically motivated object.

\begin{quote}
 Some of the standard notation and terminology of differential geometry are not present in the paper. For example, the discussion in the supplementary document from line 115 to 124 could be referred to as the volume preserving diffeomorphism of flow maps that are presented in Kambe, Tsutomu Jixin. Geometrical Theory Of Dynamical Systems And Fluid Flows (Revised Edition). Vol. 23. World Scientific Publishing Company, 2009. Also, for standard notation of differential geometry, please see Kreyszig, Erwin. Differential geometry. Courier Corporation, 2013. 
\end{quote}
As we stated in the supplementary information of the original version, the use of standard notation is avoided on purpose because, in our experience, it does map well to physical intuition. We added an additional comment to that effect in the main body of the paper on line 26, as we realized this was not made sufficiently clear. In response to the other reviewer, we have also clarified in the paper the connection between the mathematical objects $\vec{S}$, $\omega$ and $\theta$ in our paper and in the standard mathematical literature, which further highlights the mismatch between standard mathematical definitions and the physically motivated ones. For example, $\theta$ would correspond to the canonical one form in the context of symplectic geometry and the contact form in the context of contact manifolds. Both of these objects are uniquely defined mathematically. However, given that the physical object we represent with $\theta$ is a vector potential, it is not uniquely defined as it depends on the arbitrary gauge choice. Therefore the ``mathematical'' $\theta$s represent the vector potential with a specific gauge choice.

\begin{quote}
 My understanding that the geometrical representation provided is only for systems that are divergence free. (i.e., they emerge from a potential). Simultaneously they will be conservative systems.  One of the conclusions is that “Also note that the cases where the Lagrangian does not admit ……instead of Lagrangian mechanics”. This statement is not clear to the reviewer, because the authors relied on the Lagrangian- Hamiltonian relationship (i.e., Legendre transformation) in equation 10. How does equation 10 hold, and at the same time the Lagrangian would not admit a corresponding Hamiltonian? 
\end{quote}
We have greatly improved and clarified that paragraph. See lines 127-133.

\begin{quote}
 General comments and questions
 Is the study limited to divergence free vector fields only? 
\end{quote}
Yes. The study is limited to all cases where the principle of stationary action holds (i.e. yields unique solutions) which, by virtue of the Liouville theorem, corresponds to divergence free displacement vector fields.

\begin{quote}
 The standard notation of differential geometry is not quite present in the paper.
\end{quote}
This is intentional. See response above.

\begin{quote}
 Will the supplementary document be published along with the paper, or it is just for the review process? 
\end{quote}
Yes.

\begin{quote}
 Why call the Principle of Least Action the Action Principle? 
\end{quote}
We use the term in the title to not make it exceedingly long. It is not an uncommon abbreviation. For example, the book ``The action principle and partial differential equations'' by D. Christodoulou.  We generally use ``principle of stationary action'' throughout the paper.

\begin{quote}
 It is not clear how the left hand-side of the second line of equation 8 is translated to be the first variation of the integral defining the action principle. 
\end{quote}
We added the extra step that takes the contour integral and breaks it up into the two paths.

\begin{quote}
 Reference number two does not consider the PLA to be “one of the most important tools in physics”, actually, it is stating the quite opposite !
\end{quote}
Correct. It was meant more as evidence of ``its physical meaning is not completely clear'', but we agree that the location of the citation was misleading. We have moved all citations to the end of the sentence.

\begin{quote}
 Editorial comments

 Line 28, “will presents …” should be “present …”, “key point …” should be “key points …”

 In pp. 11 line 140 in the supplementary document, “Putting …” should be “Summing or Grouping …”.
\end{quote}
We thank the reviewer for taking the time to point out the typos, which have been fixed. We also have reworded the sentence.

\section{Replies to Reviewer 2}

\begin{quote}
The action principle which gives rise to the equations of motion is often obscure, framed only in mathematical terms, with little to no geometric interpretation other than the cryptic “nature chooses the path that minimizes the action” — which feels like it reveals something deep about the world, and all those “other possible paths”, but it is still mysterious.

So it is refreshing to see in this paper a concise and direct derivation of the action principle that arises from basic physical assumptions. As far as I know, this is indeed a new interpretation: I haven’t seen any other geometric construction like this regarding the action principle. I am very intrigued at the equivalence between the physical principles and the assumptions, stated in plain language, which I read in the supplement is one of the key goals of the authors’ research programme. 
\end{quote}

We thank the referee for the positive remarks.

\begin{quote}
	Now, I am curious, and would like the authors to clarify what mathematical objects are new.  The displacement field S was new to me. Is the potential $\theta$, also new?  And the counting form, lower case $\omega$, is this the symplectic 2-form familiar from Hamiltonian geometry?  If so, please state. And if not, perhaps a different symbol would be better.
\end{quote}
We added remarks when the symbols are introduced in the main text, at lines 41, 52 and 98. We also added more explanation in the supplementary material at lines 186, 204 and 380. The displacement field S is typically referred to as "the vector field" in dynamical system theory. $\theta$ roughly corresponds to the canonical one form/contact form and $\omega$ to the symplectic form, though there are caveats: we are working in an odd dimensional space so technically $\omega$ is not symplectic and $\theta$ for us is a vector potential, therefore the canonical one form/contact form correspond to a specific gauge choice of the vector potential. These are exactly the type of mismatches between the physics and the mathematics that we had to overcome before being able to connect everything. We believe the full mathematical account of these differences would clutter the exposition, so we hope the sketches we provide are sufficient.

\begin{quote}
There are some minor typos, but otherwise the language is fine and readable. In fact, the authors are gifted writers, and I appreciate that the paper is under 75 lines!
\end{quote}
We have addressed a handful of typos we found. We thank the referee for the kind words.

\begin{quote}
I have some additional comments, but not to be construed as a critique of the manuscript. As all good papers do, this one has inspired some further questions.

from line 70, “In this sense, the action principle is better understood as a feature of Hamiltonian mechanics, instead of Lagrangian mechanics.” This is remarkable for those with a classical physics education, who have read Goldstein, etc. In most books, it is Lagrangian mechanics that’s based on the action principle, and then we shoehorn a Hamiltonian into the action principle, and it never seems natural. However, the action principle in Hamiltonian form does underly the theory of canonical transformations. The authors did not deal with canonical transformations here, but is there some interesting path there to explore? What one loses is the physical interpretation of the coordinates and momenta, and I understand the authors want to maintain that connection.
\end{quote}
In the larger project we do analyze the physical motivation of the structure of phase space itself, so that we have a set of assumptions that motivate the symplectic structure, and therefore the theory of canonical transformations. What we have found is that symplectic spaces are exactly those manifolds that allow us to write densities, count of states and entropy in a frame invariant way. Canonical coordinates are essentially choices of state variables that allow those elements to be expressed in the correct units. This is why, for example, in statistical mechanics the maximization of entropy must be performed over position and momentum and not, for example, position and velocity. We have chosen not to modify the paper as we feel it would add significant length for an adequate discussion.

Some of these ideas have been explored in our papers "Hamiltonian mechanics is conservation of information entropy" and "Hamiltonian privilege". We haven't cited them as we didn't want to have too many self-citations. We are also working on a new edition of our "Assumptions of Physics" book, which we have now cited in the acknowledgments, that will contain the full account for classical systems.

\begin{quote}
In any case, I'll be paying attention to the Reverse Physics programme.
\end{quote}
We thank the reviewer for the interest in our project!

\clearpage
\bibliographystyle{apalike}
\bibliography{bibliography}
\end{document}

### 2023-06-28 Articles/2022-ActionPrinciple/ScientificReportsAmended/ActionPrincipleBody.tex
Resubmit after reviewer reports
Also changed body section of ActionPrincipleSR.tex to be consistent with ActionPrincipleBody.tex.  ActionPrincipleSR.tex additionally contains the Supplementary Information section.

@@
 We then proceed to the case of multiple independent degrees of freedom using tools from differential geometry.
-The main article will presents all the key points needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the Supplementary Information.
+The main article will present all the key points needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the Supplementary Information.
 
@@
 If assumption \ref{assum_detrev} is valid, we expect the flow of states through a closed surface to be zero: as many states flow in as flow out of the region.
-Alternatively, if we assign a probability, or probability density, to each trajectory, the assumptions requires that probability not to change, so integrating the probability over a closed surface must yield zero.
+Alternatively, if we assign a probability, or probability density, to each trajectory, the assumption requires that probability not to change, so integrating the probability over a closed surface must yield zero.
 However we see it, assumption \ref{assum_detrev} means the field is divergence-free.
-[Given that this is a three dimensional space, we can use the standard tools of vector calculus.] That is,
+[Given that this is a three-dimensional space, we can use the standard tools of vector calculus.] That is,
 \begin{equation}\label{sdof_div_free}
@@
 
-We find that a system for which \ref{assum_detrev} and \ref{assum_kineq} are valid can be characterized in terms of the action principle with a suitable Lagrangian.
-The converse is also true: if the action principle allows for a unique solution, then the conjugate momentum and the Hamiltonian are well defined and the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
+We find that a system for which \ref{assum_detrev} and \ref{assum_kineq} are valid can be characterized in terms of the principle of stationary action with a suitable Lagrangian.
+The converse is also true: if the principle of stationary action allows for a unique solution, then the conjugate momentum and the Hamiltonian are well defined and the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
 
-We have thus demystified the action principle, and turned it into a geometric property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field in phase space.
+We have thus demystified the principle of stationary action, and turned it into a geometric property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field in phase space.
 We also have a clear physical meaning: the principle of stationary action is equivalent to assuming determinism/reversibility \ref{assum_detrev} and kinematic equivalence \ref{assum_kineq}.
@@
 
-The setup to recover the action principle is the same, except that we need to substitute vector calculus with differential geometry.
+The setup to recover the principle of stationary action is the same, except that we need to substitute vector calculus with differential geometry.
 First we find that the flow through $\Sigma$ is the variation of the action using Stokes' theorem
@@
 \section*{Acknowledgments}
-This paper forms part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which
-the basic laws can be rigorously derived. The authors would like to thank Alejandro Uribe-Ahumada and Daniel Burns for helpful discussions.
+This paper forms part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+The authors would like to thank Alejandro Uribe and Daniel Burns for helpful discussions.
 

### 2023-06-28 Articles/2022-ActionPrinciple/ScientificReportsAmended/ActionPrincipleSR.tex
Resubmit after reviewer reports
Also changed body section of ActionPrincipleSR.tex to be consistent with ActionPrincipleBody.tex.  ActionPrincipleSR.tex additionally contains the Supplementary Information section.

@@
 
-While the principle of stationary action is regarded by many as one of the most important tools in physics,\cite{hamilton1834general, heaviside1903, feynman1942principle, brenier1989least, rojo2018principle} its physical meaning is not completely clear.
-First of all, the typical characterization of the Lagrangian as the difference between kinetic and potential energy fails even for simple systems, like a charged particle under a magnetic field.
+While the principle of stationary action is regarded by many as one of the most important tools in physics, its physical meaning is not completely clear.\cite{hamilton1834general, heaviside1903, feynman1942principle, brenier1989least, rojo2018principle} First of all, the typical characterization of the Lagrangian as the difference between kinetic and potential energy fails even for simple systems, like a charged particle under a magnetic field.
 Moreover, the Lagrangian for a system is not uniquely defined, making the actual value of the action for a path not directly physically significant.
@@
 The two aspects that have been sorely lacking, and that we provide, are (1) a clear geometric interpretation of the action principle and (2) a tight connection between the math and the physics it represents.
-To make the result accessible to the widest audience, we first cover the case of a single degree of freedom using standard vector calculus.
+[This has been the hardest problem to solve, as the modern mathematical concepts and notation have departed from physical intuition.
+We will therefore use, instead of the standard notation of differential geometry, an extended version of the one physicists use in general relativity.] To make the result accessible to the widest audience, we first cover the case of a single degree of freedom using standard vector calculus.
 We then proceed to the case of multiple independent degrees of freedom using tools from differential geometry.
-The main article will present all the key point needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the Supplementary Information.
+The main article will present all the key points needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the Supplementary Information.
 
@@
 that describes how states move in time.
-[Where possible, we will be writing the same expression in both vector calculus and component notations.] The time component of the displacement vector field is constrained, as we have
+[Where possible, we will be writing the same expression in both vector calculus and component notations.] In dynamical system literature, this is referred to as the vector field of the dynamical system.
+The time component of the displacement vector field is constrained, as we have
 \begin{equation}\label{sdof_time_constraint}
@@
 		- \iint_{\Sigma} \vec{S} \cdot d\vec{\Sigma} &= \iint_{\Sigma} \left( \nabla \times \vec{\theta} \right) \cdot d\vec{\Sigma} \\
-	=  \oint_{\partial \Sigma = \gamma \cup \gamma'} \vec{\theta}  \cdot d\vec{\gamma} 
+		&=  \oint_{\partial \Sigma = \gamma \cup \gamma'} \vec{\theta}  \cdot d\vec{\gamma}  \\
+		&= \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma} - \int_{\gamma'} \vec{\theta} \cdot d\vec{\gamma}' \\
 		&= \delta \int_{\gamma} \vec{\theta} \cdot d\vec{\gamma}.
@@
 
-We find that a system for which \ref{assum_detrev} and \ref{assum_kineq} are valid can be characterized in terms of the action principle with a suitable Lagrangian.
-The converse is also true: if the action principle allows for a unique solution, then the conjugate momentum and the Hamiltonian are well defined and the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
-
-We have thus demystified the action principle, and turned it into a geometric property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field in phase space.
+We find that a system for which \ref{assum_detrev} and \ref{assum_kineq} are valid can be characterized in terms of the principle of stationary action with a suitable Lagrangian.
+The converse is also true: if the principle of stationary action allows for a unique solution, then the conjugate momentum and the Hamiltonian are well defined and the system satisfies both \ref{assum_detrev} and \ref{assum_kineq}.
+
+We have thus demystified the principle of stationary action, and turned it into a geometric property: requiring the principle of stationary action is equivalent to requiring that the solutions are the field lines of a divergence-free field in phase space.
 We also have a clear physical meaning: the principle of stationary action is equivalent to assuming determinism/reversibility \ref{assum_detrev} and kinematic equivalence \ref{assum_kineq}.
@@
 Therefore if we construct any parallelepiped at equal time, two opposite sides will contain the same number of states with opposite orientation, and therefore the integral of $\omega$ over the whole surface will be zero.
+Those familiar with differential geometry will recognize $\omega$ as the symplectic form of symplectic manifolds and of the symplectic bundles of contact manifolds.
 
@@
 		\theta &= \theta_a e^a = \theta_{q^i} e^{q^i} + \theta_{p_i} e^{p_i} + \theta_t e^t \\
-		\omega &= - \partial \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
-	\end{aligned}
-\end{equation}
-The minus sign is required to match conventions. [We write the exterior derivative as $\partial \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.]
+		\omega &= - \partial \wedge \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
+	\end{aligned}
+\end{equation}
+The minus sign is required to match conventions.
+[We write the exterior derivative as $\partial \wedge \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.]
 
@@
 \section*{Data availability}
-Not applicable.
+All data generated or analysed during this study are included in this published article and its supplementary information files.
+
+\vspace{-1mm}
+\section*{Acknowledgments}
+This paper forms part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+The authors would like to thank Alejandro Uribe and Daniel Burns for helpful discussions.
 
@@
 We would find that the field lines of $\vec{B}$ are the lines and only the lines for which the integral of the vector potential is stationary.
-Similarly, we would find that the streamlines of an incompressible fluids are the lines and only the lines for which the integral of the vector potential of the flow is stationary.
+Similarly, we would find that the streamlines of an incompressible fluid are the lines and only the lines for which the integral of the vector potential of the flow is stationary.
 While these properties may be of limited practical use, the geometric insight is the same.
@@
 \end{equation*}
-The work $W$ over a path $\lambda$ can be understood as the sum of the contributions of a force field $\vec{f}$ over each infinitesimal line $\vec{d\gamma}$:
+The work $W$ over a path $\gamma$ can be understood as the sum of the contributions of a force field $\vec{f}$ over each infinitesimal line $\vec{d\gamma}$:
 \begin{equation*}
@@
 \begin{equation*}
-	F(\sigma) = \int_\sigma \omega(d\sigma)
+	F(\sigma) = \int_\sigma \omega(d\sigma),
 \end{equation*}
@@
 
-Of all generalized properties, the most impressive one is certainly the generalized Stokes' theorem. Compare the divergence (or Gauss's) theorem:
+Of all generalized properties, the most impressive one is certainly the generalized Stokes' theorem. Compare the divergence (or Gauss') theorem:
 \begin{equation*}
@@
 \end{equation*}
-This is the generalized Stokes' theorem and $\partial \wedge \omega$ is the exterior derivative of $\omega$. We use $\partial$ as we leave $d$ reserved for the infinitesimal regions.
+This is the generalized Stokes' theorem and $\partial \wedge \omega$ is the exterior derivative of $\omega$. We use $\partial \wedge$ as we leave $d$ reserved for the infinitesimal regions.
 
@@
 Note that we have broken down exactly how each assumption contributes to the mathematical setting.
-If \ref{assum_indep} fails, then $\Omega$ cannot be understood as $\omega^N$, which means we have much fewer conditions on the motion: we only know that the total flow is conserved, but we cannot break it down along each d.o.f..
+If \ref{assum_indep} fails, then $\Omega$ cannot be understood as $\omega^N$, which means we have many fewer conditions on the motion: we only know that the total flow is conserved, but we cannot break it down along each d.o.f..
 On the other hand, if \ref{assum_detrev} fails we may not have a single displacement field $\vec{S}$ and/or the form that counts states would not be closed.
@@
 In fact, the only non-trivial components of $\omega$ are components of $\vec{S}$.
-These two objects, then, are not simply related: they express the same information.
+These two objects, then, are not only related: they express the same information.
 

### 2023-06-28 Articles/2022-ActionPrinciple/ScientificReportsAmended/response1.tex
Minor edits to reviewer responses before submitting them

@@
 \end{quote}
-As we stated in the supplementary information of the original version, the use of standard notation is avoided on purpose because, in our experience, it does map well to physical intuition.
+As we stated in the Supplementary Information of the original version, the use of standard notation is avoided on purpose because, in our experience, it does not map well to physical intuition.
 We added an additional comment to that effect in the main body of the paper on line 26, as we realized this was not made sufficiently clear.
 In response to the other reviewer, we have also clarified in the paper the connection between the mathematical objects $\vec{S}$, $\omega$ and $\theta$ in our paper and in the standard mathematical literature, which further highlights the mismatch between standard mathematical definitions and the physically motivated ones.
-For example, $\theta$ would correspond to the canonical one form in the context of symplectic geometry and the contact form in the context of contact manifolds.
+For example, $\theta$ would correspond to the canonical one-form in the context of symplectic geometry and the contact form in the context of contact manifolds.
 Both of these objects are uniquely defined mathematically.
@@
  General comments and questions
+
  Is the study limited to divergence free vector fields only? 
@@
 \end{quote}
-We use the term in the title to not make it exceedingly long.
+We use the term in the title to avoid making it exceedingly long.
 It is not an uncommon abbreviation.
@@
 Christodoulou.
-We generally use ``principle of stationary action'' throughout the paper.
+We generally use ``principle of stationary action'' throughout the text of the paper.
 
@@
 \end{quote}
-We thank the reviewer for taking the time to point out the typos, which have been fixed. We also have reworded the sentence.
+We thank the reviewer for taking the time to point out the typos, which have been fixed. We also have reworded the sentence beginning ``Putting ...''.
 
@@
 We added remarks when the symbols are introduced in the main text, at lines 41, 52 and 98.
-We also added more explanation in the supplementary material at lines 186, 204 and 380.
-The displacement field S is typically referred to as "the vector field" in dynamical system theory.
-$\theta$ roughly corresponds to the canonical one form/contact form and $\omega$ to the symplectic form, though there are caveats: we are working in an odd dimensional space so technically $\omega$ is not symplectic and $\theta$ for us is a vector potential, therefore the canonical one form/contact form correspond to a specific gauge choice of the vector potential.
+We also added more explanation in the Supplementary Information at lines 186, 204 and 380.
+The displacement field S is typically referred to as ``the vector field'' in dynamical system theory.
+$\theta$ roughly corresponds to the canonical one-form/contact form and $\omega$ to the symplectic form, though there are caveats: we are working in an odd-dimensional space so technically $\omega$ is not symplectic and $\theta$ for us is a vector potential, therefore the canonical one-form/contact form correspond to a specific gauge choice of the vector potential.
 These are exactly the type of mismatches between the physics and the mathematics that we had to overcome before being able to connect everything.
@@
 \end{quote}
-We have addressed a handful of typos we found. We thank the referee for the kind words.
+We have addressed the handful of typos we found. We thank the referee for the kind words.
 
@@
 
-Some of these ideas have been explored in our papers "Hamiltonian mechanics is conservation of information entropy" and "Hamiltonian privilege".
+Some of these ideas have been explored in our papers ``Hamiltonian mechanics is conservation of information entropy'' (Studies in History and Philosophy of Modern Physics 71:60, 2020) and ``Hamiltonian privilege'' (Erkenntnis 2023).
 We haven't cited them as we didn't want to have too many self-citations.
-We are also working on a new edition of our "Assumptions of Physics" book, which we have now cited in the acknowledgments, that will contain the full account for classical systems.
+We are also working on a new edition of our \emph{Assumptions of Physics} book (Michigan Publishing 2021), which we have now cited in the acknowledgments, that will contain the full account for classical systems.
 
@@
 
-\clearpage
-\bibliographystyle{apalike}
-\bibliography{bibliography}
+%\clearpage
+%\bibliographystyle{apalike}
+%\bibliography{bibliography}
 \end{document}

### 2023-07-03 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Reviewed with Christine

@@
 That is, if it can be shown that the mathematical definition is needed to capture and only capture a particular aspect of the physical system.
-On the other hand, a mathematical definition is \textbf{unphysical} if it can be shown to require physical properties or operations that cannot be implemented physically.
+On the other hand, a mathematical definition is \textbf{unphysical} if it can be shown to require properties or operations that cannot have a physical counterpart.
 
-For example, suppose we have a set of physically distinguishable objects, meaning that we can tell them apart through an experimental test.
+For example, suppose we have a set of physically distinguishable objects, meaning that we can tell any two apart through an experimental test.
 Experimentally, the best we can do is collect finite information in finite time.
-If given an arbitrarily long time, we can collect an arbitrary amount of finite information.
+If given an arbitrarily long time, we can collect an arbitrarily large but finite amount of information.
 An element of the set%TODO: Set of what?
@@
 , then, must be identified by, at most, a countably infinite sequence of ones and zeros so that, at some point, we can tell the difference between the two sequences and, therefore, the objects.
-The set of countable sequences of zeros and one can be understood as the binary expansion of real numbers between zero and one, which has the cardinality of the continuum.
-Therefore, any definition requiring a set of physical objects to have cardinality greater than the continuum is \textbf{unphysical}.
-For example, non-separable Hilbert spaces have cardinality greater than the continuum and, therefore, would be considered unphysical under this definition.
+The set of countable sequences of zeros and ones can be understood as the binary expansion of real numbers between zero and one, which has the cardinality of the continuum.
+Therefore, any mathematical definition requiring a set of physical objects to have cardinality greater than the continuum is \textbf{unphysical}.
+For example, non-separable Hilbert spaces have cardinality greater than the continuum and, therefore, are unphysical.
 
@@
 
-\subsection{Inner vector space}
+\subsection{Inner product space}
 
@@
 That is, it is a vector space with an inner product, and its Cauchy sequences converge under the norm defined by the inner product.
-We will look at the first two structures briefly, leaving all detail in the appendix, as their physical significance is unsurprising.
+We will look at the first two structures only briefly, leaving all detail in the appendix, as their physical significance is unsurprising.
 
@@
 If one regards superpositions as physical requirements of quantum theory, one may think this structure is physical.
-However, the ability to interpret linear combinations as superpositions requires keeping track of orthonormal vectors.
-If we can only know whether two vectors are in the same ray, we cannot recover the idea of a probability amplitude, as those are the coefficients of a linear combination between normalized and orthogonal states that return another normalized state.
+However, the ability to interpret linear combinations as superpositions requires the ability to keep track of orthonormal vectors.
+If we can only know whether two vectors are in the same ray, we cannot recover the idea of probability amplitudes, as those are the coefficients of a linear combination between normalized and orthogonal states that return another normalized state.
 Therefore, the vector space structure, by itself, is not physical.
 
+%TODO: Revise paragraph to make clearer and stronger point.  
 If we add the inner product, we can define norm and orthogonality, and therefore we can indeed express superpositions.
@@
 If one is comfortable considering superpositions and the Born rule to be physical requirements of quantum mechanics, then the inner product structure is \textbf{physical}.
-However, some may not consider superpositions as representing actual physical objects but more as mathematical constructs; others may object that the Born rule, in its standard interpretation, defines the probability of transitions while a state space should be defined at equal times; others can argue that the rays define the states, and therefore the requirements should only be in terms of those objects.
-For those and other reasons, we offer an alternative justification of the physicality of inner product vector spaces.
+However, some may not consider superpositions as representing actual physical objects but more as mathematical constructs; others may object that the Born rule, in its standard interpretation, defines the probability of transitions while a state space should be defined at equal time; others can argue that the rays define the states, and therefore the requirements should only be in terms of those objects.
+For those and other reasons, we offer an alternative justification of the physicality of inner product spaces.
 Here is a summary, but the details are in the appendix.
@@
 Additionally, the inner product can be shown to be equivalent to determining the entropy of all equal mixtures of pairs of states.
-Since we must be able to prepare statistical mixtures and characterize their entropy for any physical theory and the complex inner product space structure is exactly what is required to model the correct relationships of quantum mixtures, therefore \textbf{the inner product space structure is physical}.
+These requirements are necessary in the sense that both classical and quantum mechanics need to keep track of how ensembles combine under mixing and what entropy each ensemble has.
+
+The argument can be summed up as follows.
+On physical grounds, for any physical theory, we must be able to prepare statistical mixtures and characterize their entropy.
+The complex inner product space structure is exactly required to model the correct relationships of quantum mixtures, therefore the mathematical definition is necessitated by physical requirements.
+\textbf{The inner product space structure is physical}.
 
@@
 330 Theorem 13.8	Roman, Stephen (2008), Advanced Linear Algebra, Graduate Texts in Mathematics (Third ed.), Springer, ISBN 978-0-387-72828-5}
-An inner product vector space is complete if, given a series of vectors
+An inner product space is complete if, given a series of vectors
 $$ \sum _{k=0}^{\infty }u_{k}$$
@@
 the series converges in the space.
+That is, if the series of the norm converges, the series converges.
 This characterization of completeness is in terms of linear combinations.
-It tells us that we can prepare states that are superpositions of infinitely many elements or, equivalently, as we have seen, mixed states comprised of infinitely many pure states.
+It implies that we can prepare states that are superpositions of infinitely many elements or, equivalently, as we have seen, mixed states comprised of infinitely many pure states.
 Is this a physical requirement or just mathematical convenience?
 
-Experimentally, we can create a superposition by splitting a beam along multiple paths, which is typical in quantum optics.\footnote{TODO: cite} The mathematical expression above means splitting the original state into infinitely many progressively small components and being able to place them where we please.
-However, that would assume that our experimental set-up spans an infinite region of space.
+Experimentally, we can create a superposition by splitting a beam along multiple paths, which is typical in quantum optics.\footnote{TODO: cite} The above mathematical expression means splitting the original state into infinitely many progressively smaller components, and being able to place them where we please.
+However, it also assumes that we can place them over an infinite spatial range: we are literally saying that we can prepare states that are spread over infinite regions.
 This does not sound physical.
@@
 Again, given the difficulties of understanding superposition, let us reframe the result in terms of statistical quantities.
-Suppose we can prepare pure states with arbitrarily finite expectation values for position.
+Suppose we can prepare pure states with arbitrary finite expectation values for position.
 Then, completeness means we can also prepare pure states that fall off at infinity as $\frac{1}{x^2}$.
@@
 Clearly, this does not make physical sense.
-When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that the observed value can be arbitrarily large, not that we can literally spread values over an infinite range.\footnote{This is the difference between potential infinity and actual infinity.
-Now, ``multiplying by $x$'' is not a well-defined operator in $L^2(\mathbb{R})$ unless we restrict the configuration space of our free particle to a compact interval of the real line.
-The fact that position operators must be bounded should recover some physicality, but that wouldn't be a property of the states themselves, which is what we're interested in studying.}
+When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that the observed value can be arbitrarily large, not that we can literally spread values over an infinite range.\footnote{This is the difference between potential infinity and actual infinity.}
+
+%TODO: not clear what this comment is about -  Now, ``multiplying by $x$'' is not a well-defined operator in $L^2(\mathbb{R})$ unless we restrict the configuration space of our free particle to a compact interval of the real line.
+The fact that position operators must be bounded should recover some physicality, but that wouldn't be a property of the states themselves, which is what we're interested in studying.
 
@@
 On physical grounds, we assume the existence of quantities, like position, that can take arbitrarily large values.
-This requires that, for each possible value, there must exist some state whose expectation matches that value.
-Requiring completeness mathematically forces us to posit states with an infinite expectation value, but that cannot be implemented physically.\footnote{Philosophers interested in perturbative quantum field theory have formulated interpretive strategies to make sense of some of the divergent series that arise, but surveying them would take us too far afield.
-For an example, see \cite{miller_mathematical_2023}.} The mathematical definition implicitly assumes objects that cannot be physically realized, so \textbf{the use of completeness for the space of quantum systems in unphysical.}
+This requires that, in principle, for each possible value, there must exist some state whose expectation matches that value.
+Completeness forces us to include states with an infinite expectation value, which cannot be implemented physically.
+Thus the mathematical definition implicitly assumes objects that cannot be physically realized, so \textbf{the use of completeness for the space of quantum systems is unphysical.}
 
@@
 
-It should not be surprising that closure over infinite sequences is problematic, as properly handling infinity is one of the key problems in mathematics.\footnote{Weyl stated, ``Mathematics is the science of the infinite, its goal the symbolic comprehension of the infinite with human, that is finite, means.'' Hermann Weyl, "The Open World: Three Lectures on the Metaphysical Implications of Science," (1932) as quoted in Mind and Nature: Selected Writings on Philosophy, Mathematics, and Physics (2009) ed.
-Peter Pesic.
-Several branches of mathematics were developed to handle large and small infinities.
-} We should understand the impact on the physics of using completeness for that goal.
+It should not be surprising that closure over infinite sequences is problematic, as properly handling infinity is one of the key problems in mathematics.\footnote{Taking an extreme position, Weyl stated, ``Mathematics is the science of the infinite, its goal the symbolic comprehension of the infinite with human, that is finite, means.'' Hermann Weyl, "The Open World: Three Lectures on the Metaphysical Implications of Science," (1932) as quoted in Mind and Nature: Selected Writings on Philosophy, Mathematics, and Physics (2009) ed.
+Peter Pesic.} We should understand the impact on the physics of using completeness for that goal.
 %TODO: This sentence was a bit difficult to parse.
 
-If an inner product space is finite-dimensional, it is automatically complete.\footnote{This is exactly why completeness is mathematically desirable: infinite-dimensional spaces behave like finite-dimensional ones.} If the space is not complete, we can always complete it by enlarging it.
-The prevalent attitude among physicists is that this does not pose a problem: we simply have a bigger set of objects than we strictly need, and we simply discard what we do not need.
-However, having a bigger space with redundant elements often creates bigger problems.
+If an inner product space is finite-dimensional, it is automatically complete.\footnote{This is exactly why completeness is mathematically desirable: it makes infinite-dimensional spaces behave like finite-dimensional ones.} If the space is not complete, we can always complete it by enlarging it.
+The prevalent attitude among physicists is that this does not pose a problem: we have a bigger set of objects than we strictly need, and we simply discard what we do not need.
+Unfortunately, having a bigger space can introduce new problems.
 
-For example, take the problem of defining volumes in three-dimensional Euclidean space $\mathbb{R}^3$.
-Formally, we want to find a measure $\mu(U)$ that returns the size of the region $U$.
-Naively, we may want to define it on the set of all possible subsets.
-In this case, the Banach–Tarski paradox tells us that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls of the same size.
+For example, suppose we want to define volumes in three-dimensional Euclidean space $\mathbb{R}^3$.
+Formally, we want to find a measure $\mu(U)$ that returns the volume of a region $U$.
+Naively, we could take a region to be any subset of $\mathbb{R}^3$.
+However, this leads to the Banach–Tarski paradox,\footnote{TODO: cite} which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls of the same volume as the original.
 Clearly, this does not make physical sense.
-However, we get a consistent definition if we restrict ourselves to a measure on the Borel sets instead of all sets.
+However, we get a consistent definition if we restrict regions to be Borel sets, instead of arbitrary subsets.
 Therefore, bigger is not always better in mathematics:\footnote{In another work TODO: cite, we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
@@
 The first is that two Hilbert spaces that differ by a unitary transformation are mathematically equivalent.
-Like a change of basis, a change of representation is a unitary transformation.
+For example, changes of basis and changes of variables are unitary transformations.
 The second reason is that time evolution for an isolated system is represented by a unitary transformation.
-Let us see, then, the interplay between unitary transformations and our previous findings.\footnote{More precisely, the first reason only holds if the assumptions of the Stone-von Neumann theorem hold.
-For the conditions and physical systems under which those assumptions are unwarranted and the interpretive problems of trying to understand physical equivalence for systems that are not unitarily equivalent, see \cite{ruetsche_interpreting_2011}.
-The second reason is a consequence of the Stone theorem mentioned in the introduction.
-Notice that the definition of unitary transformations requires the inner product.
-Yet another reason for its physicality!}
+The issue is that unitary transformations in a Hilbert space can map states with finite expectation values to those with infinite ones, leading to obvious problems.
 
-In a Hilbert space, we can always find a unitary transformation that maps one vector to any other.
-Therefore, there will be unitary transformations that map states with finite expectation values to infinite ones.
-This means that whether something has a finite expectation value will depend on the choice of representation.
 As a concrete example, consider the following two wave functions
@@
 \psi(x) &= \sqrt{\frac{e^{-x^2}}{\sqrt{\pi}}} \\
-\phi(x) &= \frac{1}{\sqrt{\pi(x^2 + 1)}}.
+\phi(x) &= \sqrt{\frac{1}{\pi(x^2 + 1)}}.
 \end{align}
@@
 The second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all moments above the first.
-We can now find a change of variable $y(x)$ that transforms one distribution into the other.
+We can now find a change of variable $y=y(x)$ that transforms one distribution into the other.
 We need to set
@@
 \begin{aligned}
-\int_{0}^{x} \psi^\dagger(x) \psi(x) dx &= \int_{0}^{y(x)} \phi^\dagger(x) \phi(x) dx \\
-\int_{0}^{x} \frac{e^{-x^2}}{\sqrt{\pi}} dx &= \int_{0}^{y(x)} \frac{1}{\pi(y^2 + 1)} dx \\
-\frac{\erf(x)}{2} &= \frac{\tan^{-1}(y)}{\pi} \\
-y &= \tan \left(\frac{\pi}{2}\erf(x)\right). \\
+\int_{0}^{y(x)} \phi^\dagger(y) \phi(y) dy &= \int_{0}^{x} \psi^\dagger(x) \psi(x) dx \\
+\int_{0}^{y(x)} \frac{1}{\pi(y^2 + 1)} dy &= \int_{0}^{x} \frac{e^{-x^2}}{\sqrt{\pi}} dx \\
+\frac{\tan^{-1}(y(x))}{\pi} &= \frac{\erf(x)}{2} \\
+y(x) &= \tan \left(\frac{\pi}{2}\erf(x)\right). \\
 \end{aligned}
@@
 %TODO: I worry that I'm not sure if this change of variables is not implementable by unitary operators...writing down the explicit unitary transformation would be great.
+- address in the appendix
 Therefore, the theory of Hilbert spaces, as applied to quantum mechanics, would tell us that the two representations are equivalent.
@@
 In the example above, if we increased the variance, one observer would see all the even moments change, but the other would see them all equally infinite.
-The math, as it stands, takes physically inequivalent states as equivalent, making it impossible later to separate physical problems from unphysical ones.
+The math, as it stands, takes physically inequivalent states as equivalent, making it impossible later to separate physical questions from unphysical ones.
 
@@
 We are likely missing out on useful results by not considering what is provable within the subset of physical cases.
-We, therefore, wonder whether many of the technical issues we have in quantum theory (e.
-g., the non-existence of a valid measure for path integrals, inequivalent representations for interacting theories, the non-existence of the categorical tensor product for Hilbert spaces, etc.)\footnote{See \cite[\S A.4]{glimm_quantum_1987} for the non-existence of infinite products of Lebesgue measures.
+We, therefore, wonder whether many of the technical issues we have in quantum theory (e.g.
+the non-existence of a valid measure for path integrals, inequivalent representations for interacting theories, the non-existence of the categorical tensor product for Hilbert spaces, infinite correlations, etc.)\footnote{TODO: this should be just a list of citations, possibly on each entry in the list.
+See \cite[\S A.4]{glimm_quantum_1987} for the non-existence of infinite products of Lebesgue measures.
 Glimm and Jaffe also mention a theorem by Minlos (3.4.2) suggesting that path integral measures \textit{can} be defined in Schwartz spaces that, as far as we know, has not been paid attention to in the literature.
 The locus classicus for Haag's theorem and the interaction picture in quantum field theory is still \cite{earman_haags_2006}.
-For categorical tensor products, see TODO.} having to deal with infinities and infinite correlations are simply a consequence of using unphysical objects at a fundamental level.
+For categorical tensor products, see TODO.}  are simply a consequence of using unphysical objects at a fundamental level.
 
@@
 
-A reasonable alternative would require the expectation of all polynomials of position and momentum to be finite.
+A reasonable alternative would be to require the expectation of all polynomials of position and momentum to be finite.
 In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m - P^mX^n)$ to be finite.
-This restricts our space to those functions that are infinitely smooth (i.
-e., all derivatives exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.
+This restricts our space to those functions that are infinitely smooth (i.e.
+all derivatives exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.
 
-The Schwartz space has several other interesting properties, explored in more detail in the appendix. For example:
+The Schwartz space has several other interesting properties,\cite{moretti_spectral_2017}\footnote{TODO: cite more?} explored in more detail in the appendix. For example:
 
 \begin{itemize}
-\item the expectations fully identify it - two Schwartz are the same if and only if all expectations are the same; quantum tomography on all polynomials fully reconstructs the state
-\item it is inner product space with the same norm of $L^2$
-\item it is a dense subset of $L^2$ Hilbert space - any element of the Hilbert space can be approximated by a Schwartz function with an arbitrary level of precision\footnote{Even the problematic position operators can be defined starting from Schwartz spaces \cite[Threorem 5.23]{moretti_spectral_2017}.}
+\item its elements are fully identified by the expectations - two Schwartz functions are the same if and only if all expectations are the same; quantum tomography on all polynomials fully reconstructs the state
+\item it is an inner product space with the same norm as $L^2$
+\item it is a dense subset of $L^2$ - any element of the Hilbert space can be approximated by a Schwartz function with an arbitrary level of precision
 \item it is complete with respect to the expectation values - a sequence of functions for which all expectation values of all polynomials of position and momentum converge will converge in the Schwartz space
@@
 These properties make the Schwartz space a much more reasonable candidate to capture the physics.
-The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} The point is that alternatives exist and should be fully explored.
+The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{TODO: Compare this statement with the ``weak convergence'' of von Neumann algebras.} The point is that alternatives exist and should be fully explored.
 
@@
 
-We have made the case that Hilbert spaces in quantum mechanics are \textbf{unphysical} as they necessarily bring in states with infinite expectation values, coordinate transformations that turn finite expectations into infinite ones, and unitary transformations that do the same in finite time.
-Theorems that are valid on Hilbert spaces range over these unphysical cases, meaning that we are possibly missing results and techniques that would be physically interesting.
+We have made the case that Hilbert spaces in quantum mechanics are \textbf{unphysical} as they necessarily bring in states with infinite expectation values, variable changes that turn finite expectations into infinite ones, and time evolution operators that do the same in finite time.
+This flawed characterization of infinity is the likely cause of outstanding technical problems in the foundations of quantum mechanics and quantum field theory, and makes it impossible to have a fully meaningful understanding of the theory.
 
@@
 
-This appendix provides the same arguments in the article's main body, but expanded with full mathematical detail.
+This appendix presents the same arguments in the article's main body, but expanded with full mathematical detail.
 
@@
 The first task is to show that the vector space structure is not, by itself, physical.
-Since rays of a Hilbert space represent quantum states, and rays can be defined simply on top of the vector space structure alone, one may think that the vector space structure, by itself, is physical.
+Since quantum states are represented by rays of a Hilbert space, and rays can be defined simply on top of the vector space structure alone, one may think that the vector space structure, by itself, is physical.
 The issue is that, without an inner product, one cannot make physical sense of the coefficients one uses for linear combinations, and, therefore, superpositions become ill-defined.
@@
 To show the problem, we prove the following proposition that tells us that there are infinitely many ways to decompose a vector representing a state in terms of vectors representing two other states, meaning that the coefficients of the linear combination are arbitrary.
-\begin{prop}\label{vector_insufficient}[The vector space structure is not enough to uniquely identify superpositions.] Let $\mathcal{H}$ be a complex vector space and let $\textbf{P}(\mathcal{H})$ be its projective space.
+\begin{prop}[The vector space structure is not enough to uniquely identify superpositions]\label{vector_insufficient} Let $\mathcal{H}$ be a complex vector space and let $\textbf{P}(\mathcal{H})$ be its projective space.
 Let $\psi, \phi_1, \phi_2 \in \mathcal{H}$ such that $\psi = c_1 \phi_1 + c_2 \phi_2$ for some $c_1, c_2 \in \mathbb{C}$.
 Let $\pj{\psi}, \pj{\phi_1}, \pj{\phi_2} \in \textbf{P}(\mathcal{H})$ be the corresponding rays in the projective space.
-Then for any $\hat{c}_1, \hat{c}_2 \in \mathbb{C}$ we can find $\hat{\phi}_1 \in \pj{\phi_1}$ and $\hat{\phi}_2 \in \pj{\phi_2}$ such that
+Then for any $\hat{c}_1, \hat{c}_2 \in \mathbb{C}$ we can find  $\hat{\phi}_1, \hat{\phi}_2 \in \mathcal{H}$ such that $\pj{\hat{\phi}}_1 = \pj{\phi_1}$, $\pj{\hat{\phi}}_2 = \pj{\phi_2}$ and
 $$\psi = \hat{c}_1 \hat{\phi}_1 + \hat{c}_2 \hat{\phi}_2.$$
@@
 Since they differ only by a constant, they are elements of the same ray.
-That is, $\hat{\phi}_1 \in \pj{\phi_1}$.
-Similarly, we let $\hat{\phi}_2 = \frac{c_2}{\hat{c}_2} \phi_2$ and we have $\hat{\phi}_2 \in \pj{\phi_2}$.
+That is, $\pj{\hat{\phi}}_1 = \pj{\phi_1}$.
+Similarly, we let $\hat{\phi}_2 = \frac{c_2}{\hat{c}_2} \phi_2$ and we have $\pj{\hat{\phi}}_2 = \pj{\phi_2}$.
 We have
@@
 \begin{aligned}
-\hat{c}_1 \hat{\phi}_1 + \hat{c}_2 \hat{\phi}_2 &= \frac{c_1}{\hat{c}_1} \phi_1 + \frac{c_2}{\hat{c}_2} \phi_2 \\
+\hat{c}_1 \hat{\phi}_1 + \hat{c}_2 \hat{\phi}_2 &= \hat{c}_1 \frac{c_1}{\hat{c}_1} \phi_1 + \hat{c}_2 \frac{c_2}{\hat{c}_2} \phi_2 \\
 &= c_1 \phi_1 + c_2 \phi_2 \\
@@
 The intuition that a state has a single linear decomposition in terms of other vectors requires, first of all, that the vectors are normalized.
-To identify the coefficient as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires decomposition on orthogonal vectors.\footnote{Mathematically, this means turning the vectors $\{\phi_i\}$ into an orthonormal set using Gram-Schmidt orthonormalization, which requires the inner product.} That is, linear combinations require the inner product to be understood physically as superpositions.
-Therefore, the vector space structure is insufficient to provide a clear connection to physics.
+To identify the coefficients as probability amplitudes, that is that $\sum_i |c_i|^2 = 1$, further requires decomposition on orthogonal vectors.\footnote{Mathematically, this means turning the vectors $\{\phi_i\}$ into an orthonormal set using Gram-Schmidt orthonormalization, which requires the inner product.} That is, the inner product is required for linear combinations to be understood physically as superpositions.
+Therefore, the vector space structure by itself is insufficient to provide a clear connection to physics.
 
@@
 This is an area where different interpretations may differ \cite{albert_quantum_1994, wallace_everett_2013, howard_complementarity_2021}, and the physicality of a mathematical representation should not be interpretation-dependent.
-Second, in its standard characterization, the Born rule brings in one of the most controversial aspects of the theory, measurements, which suffers from an interconnected interpretation problem.
-Lastly, the inner product is critical for both the Born rule, which is a transition probability of transition before and after the measurement, and linear decomposition into superposition, which is done at equal times.
-The goal, then, is to find a characterization of our physical requirement that is interpretation-independent, more cohesive, and perhaps more physically necessary %TODO: Not sure what ``necessary'' means here.
+Second, in its standard characterization, the Born rule brings in one of the most controversial aspects of the theory, measurements, and thus similarly suffers from an interpretation problem.
+Lastly, the inner product is critical for both the Born rule, which is a transition probability between states before and after a measurement, and linear decomposition into superposition, which is done at equal time.
+Given that we need to show that the mathematical definition captures and only captures a particular aspect of a physical system, we need to formulate a physical requirement that is interpretation-independent and grounded in experimental practice.
 
-The idea is to reformulate the requirements that lead to superposition and the Born rules in terms of equivalent requirements of statistical mixtures.
+We will show that the mathematical tools required to characterize statistical mixtures can be reformulated in terms of superpositions and the Born rule.
 Mathematically, these will be in terms of the space of density operators instead of the vector space itself.\footnote{Similar techniques can be found in the literature of Quantum Logic and of Generalized Probability Theories.
-TODO: add citations}
+TODO: add citations} Since preparation of statistical mixtures is well grounded in experimental practice, their characterization provides a well posed physical requirement.
 
@@
 The idea is that a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ exactly because there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
-Hilbert space subspaces are grouping together pure states that can provide the same statistical mixtures.
+Subspaces of a Hilbert space group together pure states that can provide equivalent statistical mixtures.
 To keep the reasoning as general as possible, we will not require $\{\phi_i\}$ to be orthogonal to each other.
@@
 Let $\{\psi_i\}, \{\phi_j\} \in \mathcal{H}$ be two finite sets of normalized vectors.
-Suppose there exists a density operator that can be expressed as a strict convex combination of either set.
+Suppose there exists a density operator that can be expressed as a strictly convex combination of either set.
 That is, we can find two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}_{>0}$ such that
@@
 \begin{proof}
-Consider $\rho = \sum_i a_i |\psi_i\>\<\psi_i|$. For a unitary vector $\varphi \in \mathcal{H}$ we have
+Consider $\rho = \sum_i a_i |\psi_i\>\<\psi_i|$. For a normalized vector $\varphi \in \mathcal{H}$ we have
 $$p(\varphi|\rho)=tr(\rho |\varphi\>\<\varphi|) = \sum_i a_i |\<\psi_i|\varphi\>|^2$$
@@
 Let $\mathcal{H}$ be a complex inner product space.
-Let $\psi \in \mathcal{H}$ be a normalized state and $\rho : \mathcal{H} \to \mathcal{H}$ a density operator of finite rank such that $\<\psi|\rho|\psi\>\neq 0$.
-Then we can express $\rho$ as a mixture of $\psi$.
-That is, there exists a finite set of normalized states $\{\psi_i\} \in \mathcal{H}$ such that $\psi_1 = \psi$ and $\rho = \sum_i p_i |\psi_i\>\<\psi_i|$.
+Let $\psi \in \mathcal{H}$ be a normalized vector and $\rho : \mathcal{H} \to \mathcal{H}$ a density operator of finite rank such that $\<\psi|\rho|\psi\>\neq 0$.
+Then we can express $\rho$ as a mixture of pure states that includes $|\psi\>\<\psi|$.
+That is, there exists a finite set of normalized vectors $\{\psi_i\} \in \mathcal{H}$ such that $\psi_1 = \psi$ and $\rho = \sum_i p_i |\psi_i\>\<\psi_i|$, where $\{p_i\} \in \mathbb{R}_{>0}$ and $\sum_i p_i = 1$.
 \end{prop}
@@
 If $\rho = |\psi\>\<\psi|$, we are done.
-If not, the two elements will identify a line, which must intersect the boundary of the convex set in two places.
+If not, the two elements $\rho$ and $|\psi\>\<\psi|$ will identify a line in $V$, which must intersect the boundary of the convex set in two places.
 One must be $\psi$ since $\psi$ is a pure state and, therefore, an extreme point.
 Let $\rho_1$ be the other point.
-If it is an extreme point, we are done as $\rho$ is an affine combination of $\psi$ and $\rho_1$.
-If not, $\rho_1$ is on a part of the boundary that is not strictly convex.
+If it is an extreme point, we are done as $\rho$ is a convex combination of $\psi$ and $\rho_1$.
+If not, $\rho_1$ is on a part of the boundary of the convex set that is not strictly convex.
 That region will be a convex set constrained on a subspace $V_1$ of $V$.
@@
 This means that we can't have $\rho_1 = \rho$ as $\<\psi|\rho|\psi\>\neq 0$.
-Therefore we have that $\rho$ is an affine combination of $\psi$ and $\rho_1$, which must be a density operator of finite rank lower than $\rho$.
+Therefore we have that $\rho$ is a strictly convex combination of $\psi$ and $\rho_1$, which must be a density operator of finite rank lower than $\rho$.
 We can express $\rho_1$ as a convex combination of finitely many extreme points $\psi_2, \psi_3, ..., \psi_n$, each of them corresponding to a pure state.
-We have that $\rho$ is an affine combination of pure states that include $\psi$.
+We thus have that $\rho$ is a strictly convex combination of pure states that includes $\psi$.
 \end{proof}
 
-\begin{prop}\label{prop_superpositionIsDecomposition}[Superposition is multiple decomposition of mixed states.]
+\begin{prop}[Superposition is non-unique decomposition of mixed states.]\label{prop_superpositionIsDecomposition}
 Let $\mathcal{H}$ be a complex vector space.
 Let $\psi \in \mathcal{H}$ be a normalized vector and $\{\phi_j\} \in \mathcal{H}$ a finite set of normalized vectors that does not contain $\psi$.
-Then $\psi$ is a superposition of $\{\phi_j\}$ if and only if there exists a density operator $\rho$ that can be expressed both as a strict convex combination of $\{\phi_i\}$ and of another set of normalized vectors that contain $\psi$.
+Then $\psi$ is a linear combination of $\{\phi_j\}$ if and only if there exists a density operator $\rho$ that can be expressed both as a strictly convex combination of $\{\phi_j\}$, and of another set of normalized vectors that contain $\psi$.
 That is, we can find a finite set of normalized vectors $\{\psi_i\} \in \mathcal{H}$ and two sequences $\{a_i\}, \{b_j\} \in \mathbb{R}_{>0}$ such that
@@
 \begin{proof}
-Suppose $\psi$ is a normalized state that can be written as $\psi = \sum_{i=1}^n c_i \phi_i$ where $\{\phi_i\} \in \mathcal{H}$ is a finite set of normalized vectors that does not contain $\psi$.
-Since $\{\phi_j\}$ are not necessarily orthogonal, it is not necessarily true that $|\<\psi|\phi_i\>| = |c_i|^2$ or that $|\<\psi|\phi_i\>| \neq 0$ for all $i$.
+Suppose $\psi$ is a normalized vector that can be written as $\psi = \sum_{i} c_i \phi_i$ where $\{\phi_i\} \in \mathcal{H}$ is a finite set of normalized vectors that does not contain $\psi$ and $\{c_i\} \in \mathbb{C}$.
+Since $\{\phi_i\}$ are not necessarily orthogonal, it is not necessarily true that $|\<\psi|\phi_i\>| = |c_i|^2$ or that $|\<\psi|\phi_i\>| \neq 0$ for all $i$.
 However, since $\psi$ is normalized and $\psi \neq \phi_i$ for all $i$, there must be at least two vectors in $\{\phi_i\}$ such that $|\<\psi|\phi_i\>| \neq 0$.
-We can then find $\{p_i\} \in \mathbb{R}+$ for which $\sum_i p_i =1$, such that $\rho = \sum_i p_i |\phi_j\>\<\phi_j|$ and $\<\psi|\rho|\psi\> \neq 0$.
-By Proposition \ref{prop_decomposition}, $\rho$ can be expressed as a convex combination of a set of normalized states that include $\psi$.
+We can then find $\{p_i\} \in \mathbb{R}_{>0}$ for which $\sum_i p_i =1$, such that $\rho = \sum_i p_i |\phi_i\>\<\phi_i|$ and $\<\psi|\rho|\psi\> \neq 0$.
+By Proposition \ref{prop_decomposition}, $\rho$ can be expressed as a strictly convex combination of a set of normalized vectors that includes $\psi$.
 
-Conversely, suppose we have a density operator $\rho$ that can be expressed as a strict convex combination of $\{\phi_i\}$ and another set of normalized vectors containing $\psi$.
-By Proposition \ref{prop_densitySpan} the span of $\{\phi_i\}$ will include $\psi$ and therefore $\psi$ is a superposition of $\{\phi_i\}$.
+Conversely, suppose we have a density operator $\rho$ that can be expressed as a strictly convex combination of $\{\phi_i\}$ and another set of normalized vectors containing $\psi$.
+By Proposition \ref{prop_densitySpan} the span of $\{\phi_i\}$ will include $\psi$ and therefore $\psi$ is a linear combination of $\{\phi_i\}$.
 \end{proof}
@@
 As for the inner product, we show that the Born rule can be understood as characterizing the entropy of equal mixtures over state pairs.
-This gives us a geometric understanding of the inner product at equal times instead of a probability of transition, which brings it more in line with the use during superposition.
+This gives us a geometric understanding of the inner product at equal time instead of as a probability of transition, which brings it more in line with the use during superposition.
 It should also be noted that the von Neumann entropy of a mixture can be understood as the minimum Shannon entropy over all possible decompositions,\footnote{TODO: cite} which again shows the connection between the geometry of the convex space of statistical ensembles and the inner product.
 
-\begin{prop}\label{prop_innerProductIsEntropy}[Inner product is an entropic structure]
-Let $\mathcal{H}$ be a complex Hilbert space. Let $I : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the von Neumann entropy of the equal mixture of two pure states. That is
+\begin{prop}[Inner product is an entropic structure]\label{prop_innerProductIsEntropy}
+Let $\mathcal{H}$ be a complex inner product space. Let $I : \mathcal{H} \times \mathcal{H} \to \mathbb{R}$ be the von Neumann entropy of the equal mixture of two pure states. That is
 \begin{equation}
@@
 
-Having shown the connection between superposition and non-unique decompositions of mixed states and between the Born rule and the entropy of mixtures, we can now use these to justify the physicality of the complex inner product space structure for quantum states.
+Having shown the connection between superposition and non-unique decompositions of mixed states, and between the Born rule and the entropy of mixtures, we can now use these to justify the physicality of the complex inner product space structure for quantum states.
 \begin{prop}
@@
 \begin{justification}
-Since physical preparations and measurements are often not perfect, a physical theory must define statistical ensembles.
-Part of that definition includes keeping track of statistical ensembles under statistical mixtures, that is, which particular mixture is obtained by mixing particular states in particular proportions.
-A particular feature of quantum mixed states is that, unlike classical states, they do not generally provide a unique decomposition in terms of pure states.
-This non-unique decomposition, as shown by proposition \ref{prop_superpositionIsDecomposition}, is equivalent to the ability to create superpositions of pure state.
+Since we can physically prepare statistical ensembles, a physical theory must be able to characterize them.
+Part of that characterization includes keeping track of which particular ensemble is obtained by mixing particular states in particular proportions.
+A defining feature of quantum mixed states is that, unlike classical mixtures, they do not provide a unique decomposition in terms of pure states.
+This non-unique decomposition, as shown by Proposition \ref{prop_superpositionIsDecomposition}, is equivalent to the ability to create superpositions of pure states.
 Therefore, the inner product structure is the exact structure needed to characterize quantum ensembles and their properties, which is required by the physics.
+Note that the complexity of the inner product space influences how many decompositions are equivalent to the same mixture, and therefore is an expression of a physical requirement.
 
-Additionally, a physical theory must be able to define each ensemble's entropy, as thermodynamics requires.
+Additionally, a physical theory must be able to define the entropy of each ensemble, as required by thermodynamics.
 In quantum mechanics, the thermodynamic entropy is recovered by the von Neumann entropy, which Proposition \ref{prop_innerProductIsEntropy} shows is exactly captured by the inner product.
-This again shows that the inner product structure is required to capture a necessary physical requirement.
+Therefore, once again, the inner product structure is required to capture a necessary physical requirement.
 \end{justification}
@@
 
-The Hellinger–Toeplitz theorem states that a symmetric operator $O$ defined on the whole Hilbert space $\mathcal{H}$ is necessarily bound.
-Therefore, an unbound operator $X$ cannot be defined on the whole space, meaning that the operator must diverge or be ill-defined on a subset of $\mathcal{H}$.
+The Hellinger–Toeplitz theorem states that a self-adjoint operator $O$ defined on the whole Hilbert space $\mathcal{H}$ is necessarily bounded.
+Therefore, an unbounded operator $X$ cannot be defined on the whole space, meaning that the operator must diverge or be ill-defined on a subset of $\mathcal{H}$.
 To see that the problem is the completeness requirement of the Hilbert space, we prove the following:
@@
 The domain $D(X)$ of the operator is a subspace of $\mathcal{H}$.
-In fact, if $||X(v_1)||_Y$ and $||X(v_2)||_Y$ are finite, so will $||X(a_1 v_1+a_2v_2)||_Y$.
+In fact, if $||X(v_1)||_Y$ and $||X(v_2)||_Y$ are finite, so is $||X(a_1 v_1+a_2v_2)||_Y$.
 Therefore $D(X)$ is an inner product space.
-It cannot be completed by the converse of the Hellinger–Toeplitz theorem.
+It cannot be complete by the converse of the Hellinger–Toeplitz theorem.
 \end{proof}
@@
 The operator is unbounded because the position can be arbitrarily large.
-We can therefore construct a Cauchy sequence $\{\psi_i\}$ such that $\lim\limits_{i \to \infty}||X\psi|| = +\infty$, that takes a sequence of vectors for which the expectation of position is larger and larger.
+We can therefore construct a Cauchy sequence $\{\psi_i\}$ such that $\lim\limits_{i \to \infty}||X\psi|| = +\infty$, that is, we can take a sequence of vectors for which the expectation of position is larger and larger.
 Because of completeness, the limit of the Cauchy sequence will be in the Hilbert space.
@@
 
-The matter would still be manageable if we had a way to keep these degenerate cases isolated, but this is not the case. Generally, a unitary operator can move vectors in and out of the domain.
+The matter would still be manageable if we had a way to keep these degenerate cases isolated, but this is not the case. In general, a unitary operator can move vectors in and out of the domain.
 \begin{prop}
@@
 That is, mathematically, two Hilbert spaces are considered indistinguishable if there exists a unitary transformation between them.
-As we saw, unitary operations can change the domain in which unbound operators are well defined: they can map states for which position is well defined onto states for which it is not.
+As we saw, unitary operations can change the domain in which unbounded operators are well defined: they can map states for which position is well defined onto states for which it is not.
 While the mathematical structure remains unchanged, the physical significance hasn't.
@@
 \end{equation}
-To find a coordinate transformation that takes finite expectations to infinite expectation, we set the initial and final wavefunctions as follow:
+To find a coordinate transformation that takes finite expectations to infinite expectations, we set the initial and final wavefunctions as follows:
 \begin{align}
@@
 \end{equation}
-The graph of the above function increase so fast that, when plotted, it looks like it has a vertical asymptote.
+The graph of the above function increases so fast that, when plotted, it looks like it has a vertical asymptote.
 Nonetheless, it is an invertible function over the whole domain and will induce a unitary transformation on the Hilbert space.\footnote{Note that we could implement a similar coordinate transformation in classical mechanics, which would also change finite moments of a distribution into infinite ones.
-It may be that, like differentiable manifolds are constrained to only differentiable coordinate transformation, we need ``statistical manifolds'' that are constrained to coordinate transformations that preserve the finiteness of moments of distributions.
+It may be that, like differentiable manifolds are constrained to only differentiable coordinate transformations, we need ``statistical manifolds'' that are constrained to coordinate transformations that preserve the finiteness of moments of distributions.
 This is a topic that requires further analysis.}
@@
 $\lim_{t \to t_0} U_t = U_{t_0}$) and respect composition in time (i.e.
-$U_t U_s = U_t+s$).
+$U_t U_s = U_{t+s}$).
 We can turn the above transformation into an evolution by setting
@@
 The idea is that the evolution will simply stretch and shrink the wavefunction intermittently, potentially oscillating expectation values from finite to infinite and vice versa in finite time.
-Clearly, this is not a meaningful time evolution, yet it will correspond to a strongly continuous one-parameter unitary group and, by Stone's theorem, will even admit a Hamiltonian.
+Clearly, this is not a physically meaningful time evolution, yet it will correspond to a strongly continuous one-parameter unitary group and, by Stone's theorem, will even admit a Hamiltonian.
 
@@
 The physicists that note that ``yes, those elements are clearly unphysical, but I don't use them, so I do not care,'' are also missing the point.
-The point is that, physically, we will want to make general statements such as ``for all observers, X is valid'' or ``under all evolution, Y is valid.'' Given that Hilbert spaces mix physically clearly meaningful objects with clearly unphysical ones, we have no way to make those statements precisely.
+The point is that, physically, we will want to make general statements such as ``for all observers, X is valid'' or ``under all evolutions, Y is valid.'' Given that Hilbert spaces mix clearly physically meaningful objects with clearly unphysical ones, we have no way to make those statements precisely.
 
-We can conclude the discussion with the following
+We can conclude the discussion with the following:
 \begin{prop}
@@
 \begin{justification}
-If a finite-dimensional space models a quantum system, and since every finite-dimensional space is complete, finite-dimensional complex Hilbert spaces are physical.
+If a finite-dimensional space models a quantum system, given that every finite-dimensional space is complete, finite-dimensional complex Hilbert spaces are physical.
 
-Some physical systems are described by unbounded discrete quantities (e.
-g., number of particles) or continuous quantities (e.
-g., position), which require infinite-dimensional spaces.
+Some physical systems are described by unbounded discrete quantities (e.g.
+number of particles) or continuous quantities (e.g.
+position), which require infinite-dimensional spaces.
 Infinite-dimensional complex inner product spaces, then, are needed to represent quantum systems.
@@
 This means that if one represents states as rays on a Hilbert space and observables as operators, the unbounded observable will not be well defined, not even as a statistical quantity, on all states.
-In particular, whether the statistical properties of an observable are well defined, whether it is possible to identify a state through tomography experimentally would be observer-dependent.
-Moreover, whether the expectation of an observer is finite or not would be something that can change in finite time.
+In particular, whether the statistical properties of an observable are well defined, or whether it is possible to identify a state through tomography experimentally, would be coordinate/representation dependent.
+Moreover, whether the expectation of an observable is finite or not would be something that can change in finite time.
 
-On physical ground, we must have that an observable has well-defined statistical properties over all states, that this property is covariant, and that it is preserved by time evolution.
+On physical grounds, we must have that an observable has well-defined statistical properties over all states, that this property is covariant, and that it is preserved by time evolution.
 Hilbert spaces do not satisfy this requirement and are, therefore, unphysical.
@@
 In other words, we need a stricter way to characterize convergence within the state space so that we can manage the infinite dimensionality.
-Conceptually, we need to treat the infinities coming from unbound quantities as potential infinities, not actual infinities: we never actually have infinitely many particles or a system positioned at infinity; we have an infinite range of possible values.
+Conceptually, we need to treat the infinities coming from unbounded quantities as potential infinities, not actual infinities: we never actually have infinitely many particles or a system positioned at infinity; we have an infinite range of possible values.
 
-A naive idea would be to require that the expectation value for all possible variables converge.
+A naive idea would be to require that the expectation value for all possible variables converges.
 Unfortunately, this doesn't work.
@@
 At a closer investigation, the Paley-Wiener theorem\footnote{TODO: find citation--Maybe Prop.
-3.115 in \cite{moretti_spectral_2017}?} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\footnote{TODO: find citation} Therefore, we cannot have compact support in both position and momentum.
-\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
+3.115 in \cite{moretti_spectral_2017}?} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\footnote{TODO: find citation} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
 Therefore the space of functions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}
@@
 The only way out, then, is to require only a certain class of observables to have finite expectation.
-Physically, it would make sense to require exactly those observables needed to reconstruct the state through quantum tomography fully.
+Physically, it would make sense to require exactly those observables needed to fully reconstruct the state through quantum tomography.
 A good set in that respect seems to be the set of polynomials of conjugate quantities, like position and momentum.
 In quantum mechanics, this means requiring all expectations of $X^nP^m = -\imath \hbar x^n\partial_x ^m$ to be finite.
-This restricts our space to those functions that are infinitely smooth (i.
-e., all derivatives exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions.
+This restricts our space to those functions that are infinitely smooth (i.e.
+all derivatives exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions.
 Requiring finite expectation for all polynomials of position and momentum, then, does not give us the usual $L^2$ Hilbert space; it gives us the Schwartz space.
 
-Not only do the expectations of all polynomials exist, but they are enough to identify the elements and define limits.
+Not only do the expectations of all polynomials exist, but they are enough to identify the elements of the space and define limits.
 The Schwartz space is, in fact, a Fréchet space \cite[Theorem V.9]{reed_methods_1980}, meaning that it is a Hausdorff space, with the topology induced by the seminorms given by the expectations, and it is complete with respect to those seminorms.
-Physically, this means that measurements on the polynomials (i.
-e., the topology of the seminorms) are enough to identify the states (i.
-e., the singletons, which are closed sets as per Hausdorff).\footnote{This is not true in classical mechanics.
-Even in the single variable case, distinct distributions have the same moments.
-However, if we assume compact support, distinct distributions have distinct moments even in the multi-variable case.\cite{moment_problem_2017} It is curious that the moment problem in quantum mechanics works differently.} Moreover, if we have a sequence for which all expectations converge, we are guaranteed to converge to another state (i.
-e., completeness with respect to the seminorms).
+Physically, this means that measurements on the polynomials (i.e.
+the topology of the seminorms) are enough to identify the states (i.e.
+the singletons, which are closed sets as per Hausdorff).\footnote{This is not true in classical mechanics.
+Even in the single variable case, distinct distributions can have the same moments.
+However, if we assume compact support, distinct distributions have distinct moments even in the multi-variable case.\cite{moment_problem_2017} It is curious that the moment problem in quantum mechanics works differently.} Moreover, if we have a sequence of states for which all expectations converge, we are guaranteed to converge to another state (i.e.
+completeness with respect to the seminorms).
 If any expectation diverges, the limit is not in the space.
@@
 To compare Schwartz and Hilbert spaces, the Schwartz space is a subspace of the $L^2$ Hilbert space taken as an inner product space.
-That is, all Schwartz functions are elements of the Hilbert space; they are closed under linear combinations and have the same inner product.
+That is, all Schwartz functions are elements of the Hilbert space: they are closed under linear combinations and have the same inner product.
 It is not complete, and it is a dense subspace, meaning that any element of the Hilbert space can always be approximated by a Schwartz function with an arbitrary level of precision.
-Or, equivalently, once a map is defined over the Schwartz space, the map is uniquely extended over the Hilbert space.
+Additionally, once a map is defined over the Schwartz space, the map is uniquely extended over the Hilbert space.
 It would then seem that all we are losing is the behavior at actual infinity.
@@
 That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property.
-Given the importance of the Fourier transform in quantum theory, this feels like an important property to have.
-Mathematically, the theory of tempered distribution, which allows us to define objects like the Dirac $\delta$-function, is built on top of the Schwartz space.
-Therefore, this space already plays a fundamental role in the definitions of the tools we already use.
+Physically, this means that we can always express a wavefunction in either the position or the momentum basis.
+Mathematically, the theory of tempered distributions, which allows us to define objects like the Dirac $\delta$-function, is built on top of the Schwartz space.
+Therefore, this space already plays a fundamental role in the definitions of the tools we already use in physics.
 
-If we restrict ourselves to the Schwartz space, then the only unitary transformations we are interested in are those that map Schwartz function to Schwartz functions.
-These are special cases of unitary transformation over Hilbert spaces and inherit all their properties, such as the existence of the inverse.
-The change of coordinate we defined in the previous section, then, is automatically ruled out.
+If we restrict ourselves to the Schwartz space, then the only unitary transformations we are interested in are those that map Schwartz functions to Schwartz functions.
+These are special cases of unitary transformations over Hilbert spaces and inherit all their properties, such as the existence of the inverse.
+The change of coordinate $y(x)$ we defined in the previous section, then, is automatically ruled out.
 It is interesting to note that two Hilbert spaces connected by such unitary transformations have essentially two separate theories of distributions as they start from two different sets of objects.
-In our mind, this shows exactly how two inequivalent physical representations can be mathematically equivalent.
-We suspect that the problem of inequivalent mathematical representations in QFT is a consequence of adding another non-physical infinity.
+This shows exactly how two inequivalent physical representations can be mathematically equivalent.
+It is not unreasonable, then, to suspect that problems of inequivalent mathematical representations in quantum field theory may be a consequence of adding a second non-physical infinity on top of the one already assumed by Hilbert spaces.
 %TODO: This last sentence is a bit controversial, and in the main text, it was looser, so I was more comfortable with conjecturing it.
@@
 Though we believe the arguments to be thoroughly convincing, we are stopping short of declaring Schwartz spaces physical.
-First, this is not the paper's goal, which is to show the unphysicality of Hilbert spaces.
-Second, we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The fact that the moment problems work differently in classical and quantum mechanics may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.\footnote{One of the issues is the fact that uniqueness works TODO}
+First, this is not the goal of the paper, which is to show the unphysicality of Hilbert spaces.
+Second, we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The fact that the moment problems work differently in classical and quantum mechanics may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.\footnote{One of the issues is the fact that uniqueness of the moment problem works differently for classical and quantum mechanics.}
 
-Now, one may be concerned that if we abandon Hilbert spaces, we would also have to abandon many convenient tools built on top.
+Now, one may be concerned that if we abandon Hilbert spaces, we would also have to abandon many convenient tools built on top of them.
 This is not the case.
-When solving an integral, we sometimes extend the domain for real to complex values, which are unphysical, because complex analysis has nicer mathematical features.
+When solving an integral, we sometimes extend the domain from real to complex values, which are unphysical, because complex analysis has nicer mathematical features.
 The result so found, however, is still a valid result in the real domain.
-Similarly, we can pose the problem on the Schwartz space, extend to the Hilbert space or the space of distributions for calculation, and then bring the result back to the Schwartz space.
-Instead, the construction of state space, especially under infinite operations, needs to be rethought.
-If we want to construct the state space for a composite system of potentially infinitely many subsystems, understanding exactly what infinities are physical and which aren't is exactly the issue.
+Similarly, we can pose a problem on the Schwartz space, extend to the Hilbert space or the space of distributions for calculation, and then bring the result back to the Schwartz space.
+What needs to be rethought, instead, are manipulations of the state spaces themselves, especially under infinite operations.
+For example, if we want to construct the state space for a composite system of arbitrarily many subsystems which are characterized by arbitrarily large observables, understanding exactly which infinities are physical and which aren't is a key issue.
 Therefore taking the Hilbert space, which we saw doesn't deal with physical infinities correctly, and making infinitely many copies of it is bound to create problems.
@@
 
-When mathematicians and philosophers noticed that mathematical theories at the time allowed to express inconsistencies (e.
-g., Russell's paradox, Berry paradox, etc.), an effort was put into strengthening the foundations of mathematics so that such paradoxes were either resolved or couldn't be expressed.
+Once mathematicians and philosophers noticed that mathematical theories allowed one to express inconsistencies (e.g.
+Russell's paradox, the Berry paradox, etc.), an effort was put into strengthening the foundations of mathematics so that such paradoxes were either resolved or couldn't be expressed.
 Concerning the mathematical foundations of physical theories, we have a similar problem: we can write statements that, though mathematically consistent, are physically inconsistent.
-Looking for and addressing these problems will ultimately lead to a solid foundation, which, we believe, is the prerequisite to solving other outstanding issues in fundamental physics.
+Looking for and addressing these problems will ultimately lead to a more solid foundation, which, we believe, is the prerequisite to solving other outstanding issues in fundamental physics.
 %TODO: We should discuss the best way to frame this!

### 2023-07-05 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Revised intro and physicality of inner product space

@@
 
-Hilbert spaces have been the cornerstone of quantum mechanics since von Neumann (\cite{von_neumann_mathematische_1996}).
-Their use made available a vast number of theorems (Gleason, Riesz, Wigner, Stone, etc.) that made up the mathematical backbone of the theory and also represented a role model of mathematical rigor that people working on quantum field theory would try to achieve decades later \cite{haag_discussion_2010}.
-Wightman even took von Neumann's book to be the fulfillment of a project Hilbert himself kicked off with his problems on the foundations of mathematics (\cite[p.
-157]{browder_hilberts_1976}): to axiomatize the physical sciences.
-But, despite the undeniable importance of wanting to set physics on a firm mathematical footing, we must remember that by ``axioms,'' they meant ``physically motivated constraints'' more than statements that can be immune to question, rethink, and, if necessary, revise.\footnote{For a more extended discussion of von Neumann's axiomatics and their physical inputs, see \cite{redei_soft_2006, landsman_axiomatization_2022}.}
+Hilbert spaces have been the cornerstone of quantum mechanics since von Neumann's 1932 book.\cite{von_neumann_mathematische_1996} While the formalism is mathematically consistent and most accept it without question, it does not have a clear physical justification.\footnote{TODO: cite?} In fact, von Neumann himself expressed his dissatisfaction not long after publication of his book, and spent considerable effort looking for alternatives.\cite{vonNeumannHilbert_1996}
 
-Now, there are well-known shortcomings of Hilbert spaces, but they are deemed unproblematic: states are rays, not vectors; the Dirac delta is a distribution, not a function; and some elements from $L^2(\mathbb{R})$ don't seem apt candidates to be wave functions.
-Given the importance Hilbert spaces have had and the issues we have let slide, it's not a surprise that they have become fossilized as parts of axiomatic systems\footnote{Even axiomatic formulations of quantum field theory \cite{haag_local_1996, horuzhy_introduction_1990} that don't need to use Hilbert spaces to make meaningful claims about quantum-mechanical observables and states can reassure physicists unfamiliar with this approach by recovering the Hilbert spaces they know and love through the GNS construction.} without even considering whether we can do quantum mechanics with weaker assumptions.
-So, despite all their success, perhaps it's time to ask if using Hilbert spaces to describe quantum states, an assumption whose physical motivation was meant to be revised as time passed, has become a dogma.
+In an effort to find a precise physical justification, we realized that no such justification can exist as infinite-dimensional Hilbert spaces hopelessly contain unphysical elements.
+To put it simply, they must include states that have infinite expectation values for position, momentum, energy and so on.
+This means that any general statement in quantum theory must be valid also on these irrelevant unphysical cases, leaving out general statements that are only valid on the more constrained and better behaved physical ones.
+We therefore suspect that some of the mathematical difficulties encountered when trying to give a more precise mathematical foundation to quantum theory stem from this fundamental mismatch.
+
+In this work we briefly present our findings, leaving more mathematical detail to the appendix.
+We will see that complex inner product spaces are indeed needed, and should be considered physical.
+We will see that the problem lies in requiring completeness, which should be understood as an unphysical mathematical requirement.
+We will show that other options exist, offering Schwartz spaces as an example.
+These guarantee the expectation values of all polynomials of position and momentum to be finite, and they are also the only space closed under Fourier transform.
 
@@
 Therefore, any mathematical definition requiring a set of physical objects to have cardinality greater than the continuum is \textbf{unphysical}.
-For example, non-separable Hilbert spaces have cardinality greater than the continuum and, therefore, are unphysical.
+For example, non-separable Hilbert spaces have cardinality greater than the continuum and, therefore, are unphysical.\footnote{TODO: cite Earman Quantum Physics in non-separable Hilbert spaces?}
 
@@
 If we add the inner product, we can define norm and orthogonality, and therefore we can indeed express superpositions.
-Additionally, the inner product structure also defines the Born rule, the transition probability during measurement.
-If one is comfortable considering superpositions and the Born rule to be physical requirements of quantum mechanics, then the inner product structure is \textbf{physical}.
-However, some may not consider superpositions as representing actual physical objects but more as mathematical constructs; others may object that the Born rule, in its standard interpretation, defines the probability of transitions while a state space should be defined at equal time; others can argue that the rays define the states, and therefore the requirements should only be in terms of those objects.
-For those and other reasons, we offer an alternative justification of the physicality of inner product spaces.
-Here is a summary, but the details are in the appendix.
+Additionally, the inner product structure defines the Born rule, the transition probability during measurement.
+While some may consider these physical requirements of quantum mechanics, their status is subject to interpretation.
+Some consider superpositions as mathematical constructs; the Born rule brings in the notion of measurements.
+We therefore reframe those requirements in terms of properties of ensembles.
+Here we provide a summary, and leave the details to the appendix.
 
-Given that, experimentally, no preparation or measurement is perfect, a physical theory must define the space of statistical ensembles.
-Mathematically, these form a convex set that physically represents a set that allows statistical mixtures.
-The geometry of the space of statistical ensembles is uniquely defined by the inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.\footnote{There is a rich literature on Generalized Probability Theory in which people reconstruct Quantum Theory by imposing requirements on convex spaces.
-TODO: cite.} For example, we can characterize a superposition in the following way: a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ if and only if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
+Since we can physically prepare statistical ensembles, a physical theory must be able to characterize them.
+Mathematically, these form a convex set, where the convex combinations are the statistical mixtures.\footnote{There is a rich literature on generalized probability theory (GPT) in which people reconstruct quantum theory by imposing requirements on convex spaces.
+TODO: cite.} In quantum mechanics, the geometry of the space of statistical ensembles is exactly defined by the complex inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.
+For example, we can characterize a superposition in the following way: a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ if and only if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
 That is, the ability to create a superposition is conceptually equivalent to the ability to prepare the same mixture in different ways, another feature particular to quantum mechanics.
@@
 These requirements are necessary in the sense that both classical and quantum mechanics need to keep track of how ensembles combine under mixing and what entropy each ensemble has.
+As such, this approach avoids the issues of interpretation.
 

### 2023-07-12 Articles/2023-HilbertQuantum/HilbertQuantum.tex
HilbertQuantum: added authors, abstract and more citations

@@
 \title{The unphysicality of Hilbert spaces}
-\author{TBD}
+\author{Gabriele Carcassi}
+\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\author{Francisco Calder\'on}
+\affiliation{Philosophy Department, University of Michigan, Ann Arbor, MI 48109}
+\author{Christine A. Aidala}
 \affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
@@
 \begin{abstract}
-TBD
+We show that Hilbert spaces should not be considered the ``correct'' spaces to represent quantum states mathematically.
+We first prove that the requirements posited by complex inner product spaces are physically justified.
+We then show that completeness in the infinite-dimensional case requires the inclusion of states with infinite expectations, coordinate transformations that take finite expectations to infinite ones and vice-verse, and time evolutions that transform finite expectations to infinite ones in finite time.
+This makes Hilbert spaces physically unsound as they model a potential infinity as an actual infinity.
+We argue that Schwartz spaces may be better candidates, as, among other things, guarantee that the expectation of all polynomials of position and momentum are finite, guarantee solution to the moment problem and are the only space closed under Fourier transform.
+We suspect that at least some problems in quantum theory related to infinities may be ultimately caused by the wrong space being used.
 \end{abstract}
@@
 
-%TODO: Revise paragraph to make clearer and stronger point.  
 If we add the inner product, we can define norm and orthogonality, and therefore we can indeed express superpositions.
@@
 Since we can physically prepare statistical ensembles, a physical theory must be able to characterize them.
-Mathematically, these form a convex set, where the convex combinations are the statistical mixtures.\footnote{There is a rich literature on generalized probability theory (GPT) in which people reconstruct quantum theory by imposing requirements on convex spaces.
-TODO: cite.} In quantum mechanics, the geometry of the space of statistical ensembles is exactly defined by the complex inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.
+Mathematically, these form a convex set, where the convex combinations are the statistical mixtures.\footnote{There is a rich literature on generalized probabilistic theory (GPT) in which people reconstruct quantum theory by imposing requirements on convex spaces.\cite{gpt_overview_2021}} In quantum mechanics, the geometry of the space of statistical ensembles is exactly defined by the complex inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.
 For example, we can characterize a superposition in the following way: a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ if and only if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
@@
 
-Experimentally, we can create a superposition by splitting a beam along multiple paths, which is typical in quantum optics.\footnote{TODO: cite} The above mathematical expression means splitting the original state into infinitely many progressively smaller components, and being able to place them where we please.
+Experimentally, we can create a superposition by splitting a beam along multiple paths, which is typical in quantum optics.\cite{banaszek2013focus} The above mathematical expression means splitting the original state into infinitely many progressively smaller components, and being able to place them where we please.
 However, it also assumes that we can place them over an infinite spatial range: we are literally saying that we can prepare states that are spread over infinite regions.
@@
 Naively, we could take a region to be any subset of $\mathbb{R}^3$.
-However, this leads to the Banach–Tarski paradox,\footnote{TODO: cite} which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls of the same volume as the original.
+However, this leads to the Banach–Tarski paradox,\cite{tao2021introduction} which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls of the same volume as the original.
 Clearly, this does not make physical sense.
 However, we get a consistent definition if we restrict regions to be Borel sets, instead of arbitrary subsets.
-Therefore, bigger is not always better in mathematics:\footnote{In another work TODO: cite, we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
+Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{AoPBook} we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
 We have often found that physical requirements naturally map to well-behaved spaces.%TODO: Is this related to the spectral theorem somehow?
@@
 The locus classicus for Haag's theorem and the interaction picture in quantum field theory is still \cite{earman_haags_2006}.
-For categorical tensor products, see TODO.}  are simply a consequence of using unphysical objects at a fundamental level.
+For categorical tensor products, see TODO.
+For now only \url{https://www-users.cse.umn.edu/~garrett/m/v/nonexistence_tensors.pdf}}  are simply a consequence of using unphysical objects at a fundamental level.
 
@@
 We will show that the mathematical tools required to characterize statistical mixtures can be reformulated in terms of superpositions and the Born rule.
-Mathematically, these will be in terms of the space of density operators instead of the vector space itself.\footnote{Similar techniques can be found in the literature of Quantum Logic and of Generalized Probability Theories.
-TODO: add citations} Since preparation of statistical mixtures is well grounded in experimental practice, their characterization provides a well posed physical requirement.
+Mathematically, these will be in terms of the space of density operators instead of the vector space itself.
+Since preparation of statistical mixtures is well grounded in experimental practice, their characterization provides a well posed physical requirement.
 
@@
 This gives us a geometric understanding of the inner product at equal time instead of as a probability of transition, which brings it more in line with the use during superposition.
-It should also be noted that the von Neumann entropy of a mixture can be understood as the minimum Shannon entropy over all possible decompositions,\footnote{TODO: cite} which again shows the connection between the geometry of the convex space of statistical ensembles and the inner product.
+It should also be noted that the von Neumann entropy of a mixture can be understood as the minimum Shannon entropy over all possible measurement contexts,\cite{nielsen_chuang_2010} which again shows the connection between the geometry of the convex space of statistical ensembles and the inner product.
 
@@
 Physically, this seems too restrictive as it excludes functions like normal distributions.
-At a closer investigation, the Paley-Wiener theorem\footnote{TODO: find citation--Maybe Prop.
-3.115 in \cite{moretti_spectral_2017}?} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\footnote{TODO: find citation} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
+At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact}\footnote{Maybe Prop.
+3.115 in \cite{moretti_spectral_2017}?} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
 Therefore the space of functions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}

### 2023-07-12 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Fixed small typo

@@
 We first prove that the requirements posited by complex inner product spaces are physically justified.
-We then show that completeness in the infinite-dimensional case requires the inclusion of states with infinite expectations, coordinate transformations that take finite expectations to infinite ones and vice-verse, and time evolutions that transform finite expectations to infinite ones in finite time.
+We then show that completeness in the infinite-dimensional case requires the inclusion of states with infinite expectations, coordinate transformations that take finite expectations to infinite ones and vice-versa, and time evolutions that transform finite expectations to infinite ones in finite time.
 This makes Hilbert spaces physically unsound as they model a potential infinity as an actual infinity.

### 2023-07-13 Articles/2023-TheWrongMath/TheWrongMath.tex
WrongMath: skeleton for possible paper

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{amsmath}
\usepackage{mathrsfs}
\usepackage{amsfonts}

\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
	colorlinks=true,
	citecolor=blue,
	urlcolor=blue,
	linkcolor=blue
}
\urlstyle{same}
\frenchspacing

\begin{document}

\title{The Wrong Math}
\author{Gabriele Carcassi, Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
It is our observations that both physicists and mathematicians do not dwell too much on the basic definitions of the mathematical tools they use. Therefore we wonder: how do we know that the mathematical structures we routinely use (e.g. Hilbert spaces, differentiable manifolds, measure theory, ...) are the ``right'' one for physics? 
In fact, what is the main feature that a mathematical structure should have to be suitable for physics? The most straight-forward answer is that a particular mathematical structure is a good match for physics if the definition corresponds to a set of physically meaningful assumptions, which then define the realm of applicability of said structure. We will see that many commonly used mathematical structures fail this test, defining a mathematical problem that is mismatched to the physical one. Our conclusion is that no significant progress can be made in fundamental physics until this situation is rectified.
\end{abstract}

\maketitle

\section{Summary}

[The whole paper in a couple of pages.]

\section{Everything you always wanted to know about Hilbert spaces (but were afraid to ask)}

Dissect Hilbert spaces and argue they are too big.

\section{The measure of a mathematical structure}

Judge by result vs judge by the definitions. No definition is always ``correct'', but some are ``wrong''.

Mathematical definitions are part of the job of the theoretical physicists. Argue that this is not the case for other tools we borrow from other fields (i.e. building codes, network standards, natural language, ...)

Proper structure (assumptions are satisfied)
Improper structure (assumptions are not satisfied)
Wrong structure (assumptions are never satistied)

Usefulness for calculation is the consequence of proper structures.

Examples: topologies as logic structures, boolean algebras for system composition.

\section{Looking for physics in all the wrong math}

Possible mistakes.

Overcomplication. If we don't have a clear idea of what we need, we are going to be pushed to use ``fancy'' tools that mathematicians/computer scientists/... like/is in their interest for us to use. Fewer things, simpler tool is better!

You model things incorrectly. Example: model the wave-function as invariant. Lack of additivity for mass.

You miss what the mathematics is capturing. Example: coordinates and units. Symplectic form and change of coordinates. Forms. Exterior derivative. Physical significance of the action principle.

\section{Finding physics}

Proper mathematical structure for generalization and comparison. Example: comparing classical and quantum mechanics.

Find the physics. Role of entropy as the definer for measures and geometry.

\section{Conclusion}

\bibliography{bibliography}

\end{document}

### 2023-07-13 Blog/TBD-QuantumStatesStableEquilibria.tex
Possible blog entry

\documentclass[aps,pra,10pt,floatfix,nofootinbib]{revtex4-2}

\usepackage{bbm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{tikz}

\newtheorem{thm}{Theorem}[section]
\newtheorem{cor}[thm]{Corollary}
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}

\theoremstyle{definition}
\newtheorem{defn}[thm]{Definition}
\newtheorem*{assump1}{Classical assumption}
\newtheorem*{assump2}{Determinism and Reversibility assumption}

\begin{document}

\section{Quantum states are stable equilibria}
TL;DR - Quantum projection processes (i.e. measurements) are stable for small perturbations.

In a previous post we showed how eigenstates, and ultimately all quantum states, can be understood as equilibria of projection processes. That is, the eigenstates are exactly those states that the projection leaves unchanged. The natural follow-up question is: what type of equilibria are they? Are they stable?

Intuitively, one may think that small perturbations will lead to a small probability of changing value. However, this is not the case. Small perturbations will return to the original value with 100\% probability. Let's see how this works.

\section{Probability change under small perturbations}

Let's take a state $\psi$, which we suppose normalized $<\psi|\psi>=1$. We now perform a small variation that leads to another normalized state. Therefore our variation is such that we have:
\begin{align*}
	\delta <\psi|\psi> &= 0 \\
	&= <\delta \psi|\psi> + <\psi|\delta \psi> \\
	<\psi|\delta \psi> &= - <\delta \psi|\psi> = - <\psi|\delta \psi>* \\
	&= \imath \epsilon
\end{align*}
where $\epsilon$ is a real number. Geometrically, we are doing a small rotation and that is why we get an imaginary number for $<\psi|\delta \psi>$.

We want to calculate how the probability changes under small variations, so let us do this first in the general case. We want to see how the probability $P(\phi|\psi)$ of finding $\phi$ given that we preapared $\psi$ changes for small variations of $\psi$ We have:
\begin{align*}
	\delta P(\phi|\psi) &= \delta <\psi|\phi><\phi|\psi> \\
	&= <\delta \psi|\phi><\phi|\psi> + <\psi|\phi><\phi|\delta \psi>
\end{align*}

In our specific case, we want to make a small variation of $\psi$ and find $\psi$ again. Therefore we substitute $\phi$ with $\psi$ and have:
\begin{align*}
	&= <\delta \psi|\psi><\psi|\psi> + <\psi|\psi><\psi|\delta \psi> \\
	&= - \imath \epsilon \cdot 1 + 1 \cdot \imath \epsilon = 0
\end{align*}
The probability of the small variation to go back to the original state is 1.

Geometrically, what happens is that, since we are making a rotation, the direction of change is perpendicular to the initial vector. The projection, therefore, to a first approximation does not change and the probability remains the same.

\section{Conclusion}

The fact that eigenstates of projections are stable equilibria highlights, again, how these processes are dissipative and non-Hamiltonian in nature. Classical Hamiltonian mechanics does not have stable equilibria or attractors, as the phase space area would necessarily have to shrink in those regions and this is prohibited by Liouville theorem.

It also reinforces the idea how quantum states are equilibria, like in thermodynamics.

\end{document}

### 2023-07-25 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Final draft after review with Christine and Francisco

@@
 \usepackage{assumptionsofphysics}
+\usepackage{tikz}
 
@@
 
-Hilbert spaces have been the cornerstone of quantum mechanics since von Neumann's 1932 book.\cite{von_neumann_mathematische_1996} While the formalism is mathematically consistent and most accept it without question, it does not have a clear physical justification.\footnote{TODO: cite?} In fact, von Neumann himself expressed his dissatisfaction not long after publication of his book, and spent considerable effort looking for alternatives.\cite{vonNeumannHilbert_1996}
+Hilbert spaces have been the cornerstone of quantum mechanics since von Neumann's 1932 book.\cite{von_neumann_mathematische_1996} While the formalism is mathematically consistent and most accept it without question, it does not have a clear physical justification.\cite{heathcote_1990, hardy_2001} In fact, von Neumann himself expressed his dissatisfaction not long after publication of his book, and spent considerable effort looking for alternatives.\cite{vonNeumannHilbert_1996}
 
@@
 
-Let us first clarify the standard we believe should be used to judge the physicality of a formal definition.
-Suppose we have a physical system under study in a given condition.
+Let us first clarify the standard we believe should be used to judge the physicality of a formal definition.\cite{redei_tension_2020, north_physics_2021} Suppose we have a physical system under study in a given condition.
 A mathematical definition is \textbf{physical} if it properly characterizes a physical system.
@@
 
-For example, suppose we have a set of physically distinguishable objects, meaning that we can tell any two apart through an experimental test.
+For example, suppose we have a set of physically distinguishable elements (e.g.
+states, bodies, regions of space), meaning that we can tell any two apart through an experimental test.
 Experimentally, the best we can do is collect finite information in finite time.
 If given an arbitrarily long time, we can collect an arbitrarily large but finite amount of information.
-An element of the set%TODO: Set of what?
-The example is clear, but this sentence is a bit unclear.
-, then, must be identified by, at most, a countably infinite sequence of ones and zeros so that, at some point, we can tell the difference between the two sequences and, therefore, the objects.
+An element of the set, then, must be identified by, at most, a countably infinite sequence of ones and zeros so that, at some point, we can tell the difference between the two sequences and, therefore, the elements.
 The set of countable sequences of zeros and ones can be understood as the binary expansion of real numbers between zero and one, which has the cardinality of the continuum.
-Therefore, any mathematical definition requiring a set of physical objects to have cardinality greater than the continuum is \textbf{unphysical}.
-For example, non-separable Hilbert spaces have cardinality greater than the continuum and, therefore, are unphysical.\footnote{TODO: cite Earman Quantum Physics in non-separable Hilbert spaces?}
+Therefore, any mathematical definition requiring a set of physical elements to have cardinality greater than the continuum is \textbf{unphysical}.
+For example, non-separable Hilbert spaces have cardinality greater than the continuum and, therefore, are unphysical.\cite{pittphilsci18363}
 
@@
 Clearly, postulating that ``the state space of quantum states has cardinality no greater than that of the continuum'' is certainly \textbf{physical}, but this tells us little.
-Let us then see what elements of Hilbert spaces we should consider physical in the context of quantum mechanics.
-%TODO: I worry that this discussion on cardinality might be confusing since the "elements of Hilbert spaces" we'll discuss are parts of the definition of a Hilbert space, not the members of a set of physical objects.
+Let us then see what features of Hilbert spaces we should consider physical in the context of quantum mechanics.
 
@@
 Additionally, the inner product structure defines the Born rule, the transition probability during measurement.
-While some may consider these physical requirements of quantum mechanics, their status is subject to interpretation.
-Some consider superpositions as mathematical constructs; the Born rule brings in the notion of measurements.
+While some may consider these physical requirements of quantum mechanics, their status is subject to interpretation.\cite{albert_quantum_1994, wallace_everett_2013, howard_complementarity_2021, ghirardi_unified_1986} Some consider superpositions as mathematical constructs; the Born rule brings in the notion of measurements.
 We therefore reframe those requirements in terms of properties of ensembles.
@@
 Since we can physically prepare statistical ensembles, a physical theory must be able to characterize them.
-Mathematically, these form a convex set, where the convex combinations are the statistical mixtures.\footnote{There is a rich literature on generalized probabilistic theory (GPT) in which people reconstruct quantum theory by imposing requirements on convex spaces.\cite{gpt_overview_2021}} In quantum mechanics, the geometry of the space of statistical ensembles is exactly defined by the complex inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.
+Mathematically, these form a convex set, where the convex combinations are the statistical mixtures.\footnote{There is a rich literature on generalized probabilistic theories (GPTs) in which people reconstruct quantum theory by imposing requirements on convex spaces.\cite{gpt_overview_2021}} In quantum mechanics, the geometry of the space of statistical ensembles is exactly defined by the complex inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.
 For example, we can characterize a superposition in the following way: a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ if and only if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
@@
 Mathematically, this means that the space includes the limit of every Cauchy sequence; physically, this property is more difficult to characterize.
-We can look at an equivalent characterization \cite[Theorem 13.8, p.
-330]{roman_2008}.
-%\footnote{Roman 2008, p.
-330 Theorem 13.8	Roman, Stephen (2008), Advanced Linear Algebra, Graduate Texts in Mathematics (Third ed.), Springer, ISBN 978-0-387-72828-5}
+We can look at an equivalent characterization \cite[Theorem 13.8]{roman_2008}.
 An inner product space is complete if, given a series of vectors
@@
 
-Experimentally, we can create a superposition by splitting a beam along multiple paths, which is typical in quantum optics.\cite{banaszek2013focus} The above mathematical expression means splitting the original state into infinitely many progressively smaller components, and being able to place them where we please.
+Experimentally, we can create a superposition by splitting a beam along multiple paths, which is typical in quantum optics.
+The above mathematical expression means splitting the original state into infinitely many progressively smaller components, and being able to place them where we please.
 However, it also assumes that we can place them over an infinite spatial range: we are literally saying that we can prepare states that are spread over infinite regions.
@@
 Clearly, this does not make physical sense.
-When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that the observed value can be arbitrarily large, not that we can literally spread values over an infinite range.\footnote{This is the difference between potential infinity and actual infinity.}
-
-%TODO: not clear what this comment is about -  Now, ``multiplying by $x$'' is not a well-defined operator in $L^2(\mathbb{R})$ unless we restrict the configuration space of our free particle to a compact interval of the real line.
-The fact that position operators must be bounded should recover some physicality, but that wouldn't be a property of the states themselves, which is what we're interested in studying.
+When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that the observed value can be arbitrarily large, not that we can literally observe and infinite value.\footnote{This is the difference between potential infinity and actual infinity.}
 
@@
 
-It should not be surprising that closure over infinite sequences is problematic, as properly handling infinity is one of the key problems in mathematics.\footnote{Taking an extreme position, Weyl stated, ``Mathematics is the science of the infinite, its goal the symbolic comprehension of the infinite with human, that is finite, means.'' Hermann Weyl, "The Open World: Three Lectures on the Metaphysical Implications of Science," (1932) as quoted in Mind and Nature: Selected Writings on Philosophy, Mathematics, and Physics (2009) ed.
-Peter Pesic.} We should understand the impact on the physics of using completeness for that goal.
-%TODO: This sentence was a bit difficult to parse.
+It should not be surprising that closure over infinite sequences is problematic, as properly handling infinity is one of the key problems in mathematics.
+We should understand how completeness tries to address that problem and its impact on the physics.
 
@@
 Naively, we could take a region to be any subset of $\mathbb{R}^3$.
-However, this leads to the Banach–Tarski paradox,\cite{tao2021introduction} which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls of the same volume as the original.
+However, this leads to the Banach–Tarski paradox,\cite{tao2021introduction} which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls each of the same volume as the original.
 Clearly, this does not make physical sense.
@@
 Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{AoPBook} we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
-We have often found that physical requirements naturally map to well-behaved spaces.%TODO: Is this related to the spectral theorem somehow?
-} we need to look more closely at what the math dragged in.
+We have often found that physical requirements naturally map to well-behaved spaces.} we need to look more closely at what the math dragged in.
 
@@
 Both states are symmetric and, therefore, will have zero expectation value for $x$.
-The first distribution is a Gaussian; therefore, all moments of the distribution, the expectation for $x^n$, are finite.
-The second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all moments above the first.
+The probability distribution for the first is a Gaussian; therefore, all moments of the distribution, the expectation for $x^n$, are finite.
+The one for the second goes to infinity like $\frac{1}{x^2}$, so the expectation will diverge for all moments above the first.
 We can now find a change of variable $y=y(x)$ that transforms one distribution into the other.
@@
 \begin{aligned}
-\int_{0}^{y(x)} \phi^\dagger(y) \phi(y) dy &= \int_{0}^{x} \psi^\dagger(x) \psi(x) dx \\
-\int_{0}^{y(x)} \frac{1}{\pi(y^2 + 1)} dy &= \int_{0}^{x} \frac{e^{-x^2}}{\sqrt{\pi}} dx \\
+\int_{0}^{y(x)} \phi^\dagger(\hat{y}) \phi(\hat{y}) d\hat{y} &= \int_{0}^{x} \psi^\dagger(\hat{x}) \psi(\hat{x}) d\hat{x} \\
+\int_{0}^{y(x)} \frac{1}{\pi(\hat{y}^2 + 1)} d\hat{y} &= \int_{0}^{x} \frac{e^{-\hat{x}^2}}{\sqrt{\pi}} d\hat{x} \\
 \frac{\tan^{-1}(y(x))}{\pi} &= \frac{\erf(x)}{2} \\
@@
 Variable changes are unitary transformations on the Hilbert space.
-%TODO: I worry that I'm not sure if this change of variables is not implementable by unitary operators...writing down the explicit unitary transformation would be great.
-- address in the appendix
 Therefore, the theory of Hilbert spaces, as applied to quantum mechanics, would tell us that the two representations are equivalent.
@@
 On physical grounds, these are clearly not equivalent.
-Reconstructing a state from its expectation values is a common technique, quantum tomography.
-If a unitary transformation can change finite expectations to infinite ones, we effectively change what can be distinguished experimentally.
-Not only can we not measure infinite values, but different finite values would get lumped into the same infinite value.
-%TODO: The comment on tomography made me think that tomographic procedures also have an uncertainty associated with them (some collaborators and I have a paper on this), so this would mean that we have a device with infinitely good resolution, which is nonsense!
-If this argument seems right, we can include it.
-In the example above, if we increased the variance, one observer would see all the even moments change, but the other would see them all equally infinite.
+Reconstructing a state from its expectation values is a common technique, quantum tomography.\cite{banaszek2013focus} If a unitary transformation can change finite expectations to infinite ones, we effectively change what can be distinguished experimentally.
+Infinite uncertainty in one frame would map to finite uncertainty in another.
+Distributions that are distinguishable in one frame would not be distinguishable in another: in the example above, if we increased the variance, one observer would see all the even moments change, but the other would see them all equally infinite.
 The math, as it stands, takes physically inequivalent states as equivalent, making it impossible later to separate physical questions from unphysical ones.
@@
 \begin{equation}
-y(t) = \cos(\omega t) x + \sin(\omega t) \tan \left(\frac{\pi}{2}\erf(x)\right).
+z(x, t) = \cos(\omega t) x + \sin(\omega t) \tan \left(\frac{\pi}{2}\erf(x)\right).
 \end{equation}
 This is clearly continuous in $t$ and, for $t=0$, we have the identity transformation, and therefore, mathematically, there is no reason why this wouldn't be a valid time evolution.
-%TODO: This would be more convincing by showing the unitary transformation reflecting the change of variables.
 What this does is stretch back and forth the wavefunction, such that the state keeps oscillating between the two functions above, making the expectation value oscillate from finite to infinite in finite time.
@@
 We, therefore, wonder whether many of the technical issues we have in quantum theory (e.g.
-the non-existence of a valid measure for path integrals, inequivalent representations for interacting theories, the non-existence of the categorical tensor product for Hilbert spaces, infinite correlations, etc.)\footnote{TODO: this should be just a list of citations, possibly on each entry in the list.
-See \cite[\S A.4]{glimm_quantum_1987} for the non-existence of infinite products of Lebesgue measures.
-Glimm and Jaffe also mention a theorem by Minlos (3.4.2) suggesting that path integral measures \textit{can} be defined in Schwartz spaces that, as far as we know, has not been paid attention to in the literature.
-The locus classicus for Haag's theorem and the interaction picture in quantum field theory is still \cite{earman_haags_2006}.
-For categorical tensor products, see TODO.
-For now only \url{https://www-users.cse.umn.edu/~garrett/m/v/nonexistence_tensors.pdf}}  are simply a consequence of using unphysical objects at a fundamental level.
+the non-existence of a valid measure for path integrals\cite{glimm_quantum_1987}, inequivalent representations for interacting theories\cite{earman_haags_2006}, the non-existence of the categorical tensor product for Hilbert spaces\cite{garrett_tensor_2010, sorkin_inside_2022}, etc.)  are simply a consequence of using unphysical objects at a fundamental level.
 
@@
 
-The Schwartz space has several other interesting properties,\cite{moretti_spectral_2017}\footnote{TODO: cite more?} explored in more detail in the appendix. For example:
+The Schwartz space has several other interesting properties,\cite{moretti_spectral_2017, reed_methods_1980} explored in more detail in the appendix. For example:
 
@@
 These properties make the Schwartz space a much more reasonable candidate to capture the physics.
-The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{TODO: Compare this statement with the ``weak convergence'' of von Neumann algebras.} The point is that alternatives exist and should be fully explored.
+The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} The point is that alternatives exist and should be fully explored.
 
@@
 We believe that meaningful progress on the foundations of physics cannot happen without understanding the implicit assumptions embedded in the most basic mathematical tools.
+
+%TODO reference formatting for 29
 
@@
 \begin{proof}
-TODO: Copy proof from other paper.
+Let $\psi, \phi \mathcal{H}$ be two normalized vectors.
+They will identify a two-dimensional subspace which can be thought, without loss of generality, as a qubit and therefore can be represented by a Bloch sphere.
+The below picture represents the intersection of the Bloch sphere with the plane identified by $\psi$ and $\phi$.
+Let $\rho = \frac{1}{2}|\psi_1\>\<\psi_1| + \frac{1}{2}|\psi_2\>\<\psi_2|$ be the density operator that represent an equal mixture of the two states.
+It will be represented by the midpoint of the chord connecting $\psi$ and $\phi$.
+Now take the line that goes through $\rho$ and the center of the sphere: $\rho$ can also be expressed as the mixture of the states $+$ and $-$ which, since they represent equal and opposite directions, form a basis.
+To diagonalize $\rho$, then, means to express it in terms of $+$ and $-$.
+
+\begin{center}
+	\begin{tikzpicture}[scale = 1]
+		\draw (0,0) circle (2);
+		\node at (-2.3,0) {$-$};
+		\node at (2.3,0) {$+$};
+		\node at (2,1.4) {$\psi$};
+		\node at (2,-1.4) {$\phi$};
+		\draw (-2,0) -- (2,0);
+		\begin{scope}
+			\clip(0,0) circle (2);
+			\draw (0,0) -- (2,1.4);
+			\draw (0,0) -- (2,-1.4);
+			\draw (1.634,1.4) -- (1.634,-1.4);
+		\end{scope}
+		\fill (1.634,0) circle (0.05);
+		\node at (1.8,.2) {$\rho$};
+	\end{tikzpicture}
+\end{center}
+
+If $\theta_{\psi\phi}$ is the angle between $\psi$ and $\phi$ on the Bloch sphere, we have
+\begin{equation}
+	|\langle \psi | \phi \rangle |^2 = \cos^2 \frac{\theta_{\psi\phi}}{2}.
+\end{equation}
+The angle is divided by two because the angle on the Bloch sphere (i.e.
+in physical space) is double the angle in the Hilbert space.
+For example, the angle between $z^+$ and $z^-$ on the Bloch sphere is $\pi$, but the inner product between them is zero (i.e.
+opposite directions in physical space correspond to orthogonal states) and therefore the angle in the Hilbert space is $\pi/2$.
+
+Now we express $\psi$ and $\phi$ in terms of $+$ and $-$, remembering that they form a basis.
+Given that $\rho$ is at the midpoint, the figure is vertically symmetric.
+The angle between $\psi$ and $+$, then, is half of $\theta_{\psi\phi}$.
+The inner product between $\psi$ and $+$ is
+\begin{equation}
+	\begin{aligned}
+		|\langle \psi | + \rangle |^2 &= \cos^2 \frac{\theta_{\psi +}}{2} \\
+		&= \cos^2 \frac{\theta_{\psi\phi}}{4}.
+	\end{aligned}
+\end{equation}
+Keeping in mind that we are composing vectors in the Hilbert space (and not in the geometry of the physical space) we have
+\begin{align*}
+	\left|\psi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>+\sin\frac{\theta_{\psi\phi}}{4}\left|-\right> \\
+	\left|\phi\right>&=\cos\frac{\theta_{\psi\phi}}{4}\left|+\right>-\sin\frac{\theta_{\psi\phi}}{4}\left|-\right>.
+\end{align*}
+
+The density matrices corresponding to the pure states are
+\begin{align*}
+	\left|\psi\right>\left<\psi\right|&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right|\\
+	&+\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
+	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right| \\
+	\left|\phi\right>\left<\phi\right|&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right|\\
+	&-\cos\frac{\theta_{\psi\phi}}{4}\sin\frac{\theta_{\psi\phi}}{4}\left(\left|+\right>\left<-\right|+\left|-\right>\left<+\right|\right) \\
+	&+\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right|.
+\end{align*}
+
+We can now calculate the mixture
+\begin{align*}
+	\frac{1}{2}(|\psi\rangle\langle\psi| &+ |\phi\rangle\langle\phi|) \\
+	&=\cos^2\frac{\theta_{\psi\phi}}{4}\left|+\right>\left<+\right| +\sin^2\frac{\theta_{\psi\phi}}{4}\left|-\right>\left<-\right| \\
+	&=\frac{1+\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|+\right>\left<+\right| +\frac{1-\cos\frac{\theta_{\psi\phi}}{2}}{2}\left|-\right>\left<-\right| \\
+	&=\frac{1+|\langle\psi|\phi\rangle|}{2}\left|+\right>\left<+\right| +\frac{1-|\langle\psi|\phi\rangle|}{2}\left|-\right>\left<-\right|. \\
+\end{align*}
+
+As $\rho$ is in a diagonal form, the entropy is given by
+\begin{equation}\label{entropy}
+	H(\rho) = H\left(\frac{1+|\langle\psi|\phi\rangle|}{2}, \frac{1-|\langle\psi|\phi\rangle|}{2}\right).
+\end{equation}
+If we define $p(\psi, \phi) = |\langle\psi|\phi\rangle|$ and $I(\psi, \phi) = H( \frac{1}{2}|\psi_1\>\<\psi_1| + \frac{1}{2}|\psi_2\>\<\psi_2|)$ we have
+\begin{align}\label{entropy}
+	f(p) & = - \frac{1+p}{2} \log \frac{1+p}{2} 
+	- \frac{1-p}{2} \log \frac{1-p}{2} \\
+	I(\psi, \phi) &= f(p(\psi, \phi)).
+\end{align}
+Note that $f$ restricted to the domain $[0,1]$, the domain on which probability is defined, is invertible.
 \end{proof}
@@
 \begin{aligned}
-\int_{0}^{y(x)} \frac{1}{\pi(y^2 + 1)} dy &= \int_{0}^{x} \frac{e^{-x^2}}{\sqrt{\pi}} dx \\
+\int_{0}^{y(x)} \frac{1}{\pi(\hat{y}^2 + 1)} d\hat{y} &= \int_{0}^{x} \frac{e^{-\hat{x}^2}}{\sqrt{\pi}} d\hat{x} \\
 \frac{\tan^{-1}(y(x))}{\pi} &= \frac{\erf(x)}{2} \\
@@
 \begin{equation}
-y(x, t) = \cos(\omega t) x + \sin(\omega t) \tan \left(\frac{\pi}{2}\erf(x)\right).
+z(x, t) = \cos(\omega t) x + \sin(\omega t) \tan \left(\frac{\pi}{2}\erf(x)\right).
 \end{equation}
@@
 This shows exactly how two inequivalent physical representations can be mathematically equivalent.
-It is not unreasonable, then, to suspect that problems of inequivalent mathematical representations in quantum field theory may be a consequence of adding a second non-physical infinity on top of the one already assumed by Hilbert spaces.
-%TODO: This last sentence is a bit controversial, and in the main text, it was looser, so I was more comfortable with conjecturing it.
-In the end, I leave it up to you two!
 
@@
 Looking for and addressing these problems will ultimately lead to a more solid foundation, which, we believe, is the prerequisite to solving other outstanding issues in fundamental physics.
-%TODO: We should discuss the best way to frame this!
 

### 2023-07-25 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Fixed small typos in the proof and added two minor TODOs

@@
 \begin{proof}
-Let $\psi, \phi \mathcal{H}$ be two normalized vectors.
+Let $\psi, \phi \in \mathcal{H}$ be two normalized vectors.
 They will identify a two-dimensional subspace which can be thought, without loss of generality, as a qubit and therefore can be represented by a Bloch sphere.
 The below picture represents the intersection of the Bloch sphere with the plane identified by $\psi$ and $\phi$.
-Let $\rho = \frac{1}{2}|\psi_1\>\<\psi_1| + \frac{1}{2}|\psi_2\>\<\psi_2|$ be the density operator that represent an equal mixture of the two states.
+Let $\rho = \frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\phi\>\<\phi|$ be the density operator representing an equal mixture of the two states.
 It will be represented by the midpoint of the chord connecting $\psi$ and $\phi$.
@@
 in physical space) is double the angle in the Hilbert space.
-For example, the angle between $z^+$ and $z^-$ on the Bloch sphere is $\pi$, but the inner product between them is zero (i.e.
+For example, the angle between $+$ and $-$ on the Bloch sphere is $\pi$, but the inner product between them is zero (i.e.
 opposite directions in physical space correspond to orthogonal states) and therefore the angle in the Hilbert space is $\pi/2$.
@@
 \end{equation}
-If we define $p(\psi, \phi) = |\langle\psi|\phi\rangle|$ and $I(\psi, \phi) = H( \frac{1}{2}|\psi_1\>\<\psi_1| + \frac{1}{2}|\psi_2\>\<\psi_2|)$ we have
+If we define $p(\psi, \phi) = |\langle\psi|\phi\rangle|$ and $I(\psi, \phi) = H( \frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\phi\>\<\phi|)$ we have %TODO: Define f.
+It looks like a von Neumann entropy, but we haven't used the notation f(p) before.
 \begin{align}\label{entropy}
@@
 Physically, this seems too restrictive as it excludes functions like normal distributions.
-At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact}\footnote{Maybe Prop.
+At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact}\footnote{%TODO: Delete this footnote?
+Maybe Prop.
 3.115 in \cite{moretti_spectral_2017}?} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.

### 2023-07-26 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Fixed minor issues in the bibliography and resolved minor issues in the appendix

@@
 However, we get a consistent definition if we restrict regions to be Borel sets, instead of arbitrary subsets.
-Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{AoPBook} we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
+Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{aop-book} we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
 We have often found that physical requirements naturally map to well-behaved spaces.} we need to look more closely at what the math dragged in.
@@
 We believe that meaningful progress on the foundations of physics cannot happen without understanding the implicit assumptions embedded in the most basic mathematical tools.
-
-%TODO reference formatting for 29
 
@@
 \begin{equation}
-I(\psi_1, \psi_2) = S\left(\frac{1}{2}|\psi_1\>\<\psi_1| + \frac{1}{2}|\psi_2\>\<\psi_2|\right).
+I(\psi, \phi) = S\left(\frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\phi\>\<\phi|\right).
 \end{equation}
@@
 \begin{equation}
-p(\psi_1, \psi_2) = |\<\psi_1| \psi_2\>|^2.
+p(\psi, \phi) = |\<\psi| \phi\>|^2.
 \end{equation}
@@
 \begin{equation}
-I(\psi_1, \psi_2) = f(p(\psi_1, \psi_2)).
+I(\psi, \phi) = f(p(\psi, \phi)).
 \end{equation}
@@
 \end{equation}
-If we define $p(\psi, \phi) = |\langle\psi|\phi\rangle|$ and $I(\psi, \phi) = H( \frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\phi\>\<\phi|)$ we have %TODO: Define f.
-It looks like a von Neumann entropy, but we haven't used the notation f(p) before.
+If we define $p(\psi, \phi) = |\langle\psi|\phi\rangle|$ and $I(\psi, \phi) = H( \frac{1}{2}|\psi\>\<\psi| + \frac{1}{2}|\phi\>\<\phi|)$ we have
+\begin{align}\label{entropy}
+	I(\psi, \phi) &= f(p(\psi, \phi))
+\end{align}
+where
 \begin{align}\label{entropy}
 	f(p) & = - \frac{1+p}{2} \log \frac{1+p}{2} 
-	- \frac{1-p}{2} \log \frac{1-p}{2} \\
-	I(\psi, \phi) &= f(p(\psi, \phi)).
+	- \frac{1-p}{2} \log \frac{1-p}{2}.
 \end{align}
-Note that $f$ restricted to the domain $[0,1]$, the domain on which probability is defined, is invertible.
+Note that $f$ restricted to the domain $[0,1]$, the domain on which probability is defined, is invertible. We have thus proved the proposition.
 \end{proof}
@@
 Physically, this seems too restrictive as it excludes functions like normal distributions.
-At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact}\footnote{%TODO: Delete this footnote?
-Maybe Prop.
-3.115 in \cite{moretti_spectral_2017}?} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
+At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
 Therefore the space of functions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}

### 2023-08-11 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Incorporating feedback

@@
 This makes Hilbert spaces physically unsound as they model a potential infinity as an actual infinity.
-We argue that Schwartz spaces may be better candidates, as, among other things, guarantee that the expectation of all polynomials of position and momentum are finite, guarantee solution to the moment problem and are the only space closed under Fourier transform.
 We suspect that at least some problems in quantum theory related to infinities may be ultimately caused by the wrong space being used.
+We strongly believe a better solution can be found, and we look at Schwartz spaces for inspiration, as, among other things, they guarantee that the expectation of all polynomials of position and momentum are finite, guarantee solution to the moment problem and are the only space closed under Fourier transform.
 \end{abstract}
@@
 To put it simply, they must include states that have infinite expectation values for position, momentum, energy and so on.
-This means that any general statement in quantum theory must be valid also on these irrelevant unphysical cases, leaving out general statements that are only valid on the more constrained and better behaved physical ones.
+This means that a truly general statement in quantum theory must be valid also on these irrelevant unphysical cases, with the consequence that either physical results are left out, or they are obscured in a complicated patchwork of theorems with different domains of validity.
 We therefore suspect that some of the mathematical difficulties encountered when trying to give a more precise mathematical foundation to quantum theory stem from this fundamental mismatch.
@@
 That is, the ability to create a superposition is conceptually equivalent to the ability to prepare the same mixture in different ways, another feature particular to quantum mechanics.
-Additionally, the inner product can be shown to be equivalent to determining the entropy of all equal mixtures of pairs of states.
+Additionally, the norm of the inner product, together with relative phases, can be shown to be equivalent to determining the entropy of all equal mixtures of pairs of states.
 These requirements are necessary in the sense that both classical and quantum mechanics need to keep track of how ensembles combine under mixing and what entropy each ensemble has.
@@
 While we do need infinite-dimensional spaces to handle unbound quantities, such as the number of particles, or continuous quantities, such as position, it seems that Cauchy-completeness is not the proper characterization.
-While finding the proper one goes beyond the scope of this work, we want to show that alternatives are possible.
+While finding the proper one goes beyond the scope of this work, we want to show that fruitful lines of inquiry do exist.
 
@@
 These properties make the Schwartz space a much more reasonable candidate to capture the physics.
-The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} The point is that alternatives exist and should be fully explored.
+The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} To again be clear, we are not proposing this as a solution.
+We are merely pointing out that alternatives do exist and should be fully explored.\footnote{Note that $C^*$-algebras model observables and therefore do not solve the problem of modeling at hand.}
 
@@
 As for the inner product, we show that the Born rule can be understood as characterizing the entropy of equal mixtures over state pairs.
-This gives us a geometric understanding of the inner product at equal time instead of as a probability of transition, which brings it more in line with the use during superposition.
+Note that the Born rule does not depend on the absolute phase, but it does depend on relative phases.
+If one expands a vector using superposition, in fact, the relative phase will introduce interference terms.
+Therefore recovering the Born rule means recovering the inner product up to total phases.
+
+Associating the Born rule with entropy gives us a geometric understanding of the inner product at equal time instead of as a probability of transition, which brings it more in line with the use during superposition.
 It should also be noted that the von Neumann entropy of a mixture can be understood as the minimum Shannon entropy over all possible measurement contexts,\cite{nielsen_chuang_2010} which again shows the connection between the geometry of the convex space of statistical ensembles and the inner product.
@@
 
-More concretely, a change of variable $X$ to $Y$ would be a unitary operator as it is just a change of representation.
-However, we can construct changes of variables such that the domain of $X$ is not exactly mapped to the domain of $Y$.
+More concretely, a change of variable $x$ to $y$ induces a unitary operator as it is just a change of representation.
+The transformation rules for wavefunctions simply follow by the fact that $\psi^\dagger(x) \psi(x) = \rho(x)$ gives us a probability density over $x$.
+Therefore we must have
+\begin{equation}
+	\begin{aligned}
+		\rho(x) dx &= \rho(y) dy \\
+		\rho(y) &= \rho(x) \frac{dx}{dy} \\
+		\psi^\dagger(y) \psi(y) &= \psi^\dagger(x) \psi(x) \frac{dx}{dy} \\
+		\psi(y) &= \psi(x) \sqrt{\frac{dx}{dy}}.
+	\end{aligned}
+\end{equation}
+This is a unitary transformation.
+
+\begin{prop}
+	Let $\mathcal{H} = L^2{\mathbb{R}}$ be the space of square integrable functions.
+Let $f : \mathbb{R} \to \mathbb{R}$ be an orientation-preserving diffeomorphism.
+Then the map $U : \mathcal{H} \to \mathcal{H}$ such that
+	$\phi(y) = U(\psi(x)) \mapsto \psi(x(y)) \sqrt{\left|\frac{df}{dx}\right|^{-1}}$ is a unitary transformation.
+\end{prop}
+\begin{proof}
+	We have:
+	\begin{equation}
+		\begin{aligned}
+			\int \phi^\dagger(y) \phi(y) dy &= \int U(\psi(x))^\dagger U(\psi(x)) dy \\
+			 = \int \psi^\dagger(x(y)) &\sqrt{\left|\frac{df}{dx}\right|^{-1}}\psi(x(y)) \sqrt{\left|\frac{df}{dx}\right|^{-1}} dy \\
+			 &= \int \psi^\dagger(x(y)) \psi(x(y)) \left|\frac{df}{dx}\right|^{-1} dy \\
+			 &= \int \psi^\dagger(x) \psi(x) dx.
+		\end{aligned}
+	\end{equation}
+	The map $U$ preserved the inner product and therefore is a unitary transformation.
+\end{proof}
+
+The issue is that the position operator $X$ and the position operator $Y$ are different operators, and the domain of $X$ will be in general different from the domain of $Y$.
 Let $y=y(x)$ be our change of variable such that $y(0) = 0$.

### 2023-08-12 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Added comments from Christine and Francisco

@@
 That is, the ability to create a superposition is conceptually equivalent to the ability to prepare the same mixture in different ways, another feature particular to quantum mechanics.
-Additionally, the norm of the inner product, together with relative phases, can be shown to be equivalent to determining the entropy of all equal mixtures of pairs of states.
+Additionally, the norm of the inner product (i.e.
+the probabilities from the Born rule), together with relative phases, can be shown to be equivalent to determining the entropy of all equal mixtures of pairs of states.
 These requirements are necessary in the sense that both classical and quantum mechanics need to keep track of how ensembles combine under mixing and what entropy each ensemble has.
@@
 
-The Schwartz space has several other interesting properties,\cite{moretti_spectral_2017, reed_methods_1980} explored in more detail in the appendix. For example:
+The Schwartz space has several other interesting properties,\cite{moretti_spectral_2017, reed_methods_1980, hall_quantum_2013} explored in more detail in the appendix. For example:
 
@@
 \item it plays a fundamental role in the theory of distributions - objects like the $\delta$-function are mathematically defined on top of the Schwartz space
+\item the Hermite polynomials, the solutions of the harmonic oscillator, are Schwartz functions
 \end{itemize}
@@
 The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} To again be clear, we are not proposing this as a solution.
-We are merely pointing out that alternatives do exist and should be fully explored.\footnote{Note that $C^*$-algebras model observables and therefore do not solve the problem of modeling at hand.}
+We are merely pointing out that alternatives do exist and should be fully explored.\footnote{Note that $C^*$-algebras model observables and therefore do not solve the problem at hand of modeling states.}
 
@@
 If one expands a vector using superposition, in fact, the relative phase will introduce interference terms.
-Therefore recovering the Born rule means recovering the inner product up to total phases.
+Therefore recovering the Born rule means recovering the inner product up to arbitrary absolute phases, which do not contain any physical content.
 
@@
 		\rho(y) &= \rho(x) \frac{dx}{dy} \\
-		\psi^\dagger(y) \psi(y) &= \psi^\dagger(x) \psi(x) \frac{dx}{dy} \\
-		\psi(y) &= \psi(x) \sqrt{\frac{dx}{dy}}.
+		\psi^\dagger(y) \psi(y) &= \psi^\dagger(x) \psi(x) \frac{dx}{dy}.
 	\end{aligned}
 \end{equation}
-This is a unitary transformation.
+Therefore we can set 
+\begin{equation}
+	\begin{aligned}
+		\psi(y) &= \psi(x(y)) \sqrt{\frac{dx}{dy}}.
+	\end{aligned}
+\end{equation}
+If the change of variable does not change orientation, the above transformation is unitary.\footnote{If the orientation changes, the transformation is anti-unitary.}
 
 \begin{prop}
-	Let $\mathcal{H} = L^2{\mathbb{R}}$ be the space of square integrable functions.
+	Let $\mathcal{H} = L^2(\mathbb{R})$ be the space of square integrable functions.
 Let $f : \mathbb{R} \to \mathbb{R}$ be an orientation-preserving diffeomorphism.
 Then the map $U : \mathcal{H} \to \mathcal{H}$ such that
-	$\phi(y) = U(\psi(x)) \mapsto \psi(x(y)) \sqrt{\left|\frac{df}{dx}\right|^{-1}}$ is a unitary transformation.
+	$\phi(y) = U(\psi(x)) \mapsto \psi(f^{-1}(y)) \sqrt{\left(\frac{df}{dx}\right)^{-1}}$ is a unitary transformation.
 \end{prop}
 \begin{proof}
+	Note that $f$ models a change of variables, therefore, in physics notation, $y = y(x) = f(x)$ and $x = x(y) = f^{-1}(y)$.
+Also note that both variables are real, therefore the Jacobian $\frac{df}{dx}$, which exists and  is nonzero because $f$ is a diffeomorphism, is real valued.
+Therefore $\left(\frac{df}{dx}\right)^\dagger = \frac{df}{dx}$.
 We have:
@@
 			\int \phi^\dagger(y) \phi(y) dy &= \int U(\psi(x))^\dagger U(\psi(x)) dy \\
-			 = \int \psi^\dagger(x(y)) &\sqrt{\left|\frac{df}{dx}\right|^{-1}}\psi(x(y)) \sqrt{\left|\frac{df}{dx}\right|^{-1}} dy \\
-			 &= \int \psi^\dagger(x(y)) \psi(x(y)) \left|\frac{df}{dx}\right|^{-1} dy \\
+			 = \int \psi^\dagger(f^{-1}(y)) &\sqrt{\left(\frac{df}{dx}\right)^{-1}}\psi(f^{-1}(y)) \sqrt{\left(\frac{df}{dx}\right)^{-1}} dy \\
+			 &= \int \psi^\dagger(f^{-1}(y)) \psi(f^{-1}(y)) \left(\frac{df}{dx}\right)^{-1} dy \\
+			 &= \int \psi^\dagger(x(y)) \psi(x(y)) \frac{dx}{dy} dy \\
 			 &= \int \psi^\dagger(x) \psi(x) dx.

### 2023-08-12 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Added last reference

@@
 
-The Schwartz space has several other interesting properties,\cite{moretti_spectral_2017, reed_methods_1980, hall_quantum_2013} explored in more detail in the appendix. For example:
+The Schwartz space has several other interesting properties,\cite{moretti_spectral_2017, reed_methods_1980, hall_quantum_2013, rauch_solution_1991} explored in more detail in the appendix.
+For example:
 

### 2023-08-14 Articles/2022-ActionPrinciple/ActionPrincipleArchive.tex
ActionPrinciple: updating archive version

@@
 \documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
-%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
-%\documentclass[10pt,twocolumn,letterpaper]{article}
 
@@
 The two aspects that have been sorely lacking, and that we provide, are (1) a clear geometric interpretation of the action principle and (2) a tight connection between the math and the physics it represents.
-To make the result accessible to the widest audience, we first cover the case of a single degree of freedom using standard vector calculus.
+[This has been the hardest problem to solve, as the modern mathematical concepts and notation have departed from physical intuition.
+We will therefore use, instead of the standard notation of differential geometry, an extended version of the one physicists use in general relativity.] To make the result accessible to the widest audience, we first cover the case of a single degree of freedom using standard vector calculus.
 We then proceed to the case of multiple independent degrees of freedom using tools from differential geometry.
-The main article will presents all the key point needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the appendix.
+The main article will present all the key points needed to follow the argument and its physical and geometric meaning, leaving the mathematical details and calculations to the Supplementary Information.
 
@@
 \end{equation}
-that describes how states move in time.\footnote{Where possible, we will be writing the same expression in both vector calculus and component notations.} The time component of the displacement vector field is constrained, as we have
+that describes how states move in time.\footnote{Where possible, we will be writing the same expression in both vector calculus and component notations.} In dynamical system literature, this is referred to as the vector field of the dynamical system.
+The time component of the displacement vector field is constrained, as we have
 \begin{equation}\label{sdof_time_constraint}
@@
 If assumption \ref{assum_detrev} is valid, we expect the flow of states through a closed surface to be zero: as many states flow in as flow out of the region.
-Alternatively, if we assign a probability, or probability density, to each trajectory, the assumptions requires that probability not to change, so integrating the probability over a closed surface must yield zero.
-However we see it, assumption \ref{assum_detrev} means the field is divergence-free,\footnote{Given that this is a three dimensional space, we can use the standard tools of vector calculus.} that is
+Alternatively, if we assign a probability, or probability density, to each trajectory, the assumption requires that probability not to change, so integrating the probability over a closed surface must yield zero.
+However we see it, assumption \ref{assum_detrev} means the field is divergence-free.\footnote{Given that this is a three-dimensional space, we can use the standard tools of vector calculus.} That is,
 \begin{equation}\label{sdof_div_free}
@@
 Therefore if we construct any parallelepiped at equal time, two opposite sides will contain the same number of states with opposite orientation, and therefore the integral of $\omega$ over the whole surface will be zero.
-
-Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative\footnote{We write the exterior derivative as $\partial \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.} of a covector $\theta$:
+Those familiar with differential geometry will recognize $\omega$ as the symplectic form of symplectic manifolds and of the symplectic bundles of contact manifolds.
+
+Since $\omega$ is closed, in every contractible region it can be expressed as the exterior derivative of a covector $\theta$:
 \begin{equation}\label{mdof_form_potential}
@@
 		\theta &= \theta_a e^a = \theta_{q^i} e^{q^i} + \theta_{p_i} e^{p_i} + \theta_t e^t \\
-		\omega &= - \partial \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
-	\end{aligned}
-\end{equation}
-The minus sign is required to match conventions.
+		\omega &= - \partial \wedge \theta = - \left( \partial_a \theta_b - \partial_b \theta_a \right) e^a \otimes e^b
+	\end{aligned}
+\end{equation}
+The minus sign is required to match conventions.\footnote{We write the exterior derivative as $\partial \wedge \omega$ instead of $d \omega$ so that $d$ is only used for infinitesimal line or surface elements.}
 
@@
 
+\vspace{-1mm}
+\section*{Acknowledgments}
+This paper forms part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+The authors would like to thank Alejandro Uribe and Daniel Burns for helpful discussions.
+
 \bibliography{bibliography}
 
-\section*{Appendix}
-
-In this section we will review the arguments in more detail.
+\section*{Supplementary Information}
+
+In this Supplementary Information we will review the arguments in more detail.
 Given that this article is still aimed at a physics audience, the focus will be conceptual rigor, rather than mathematical rigor.
@@
 
-Now, since every functional has a corresponding form, $\partial F$ must have a corresponding $k+1$-form which must be fully identified by $\omega$. We note it as $\partial \omega$ and write
-\begin{equation*}
-	\partial F(\Sigma) = \int_\Sigma \partial \omega(d\Sigma) = \oint_{\sigma = \partial \Sigma} \omega(d\sigma) = F(\partial \Sigma).
-\end{equation*}
-This is the generalized Stokes' theorem and $\partial \omega$ is the exterior derivative of $\omega$. We use $\partial$ as we leave $d$ reserved for the infinitesimal regions.
+Now, since every functional has a corresponding form, $\partial F$ must have a corresponding $k+1$-form which must be fully identified by $\omega$. We note it as $\partial \wedge \omega$ and write
+\begin{equation*}
+	\partial F(\Sigma) = \int_\Sigma \partial \wedge \omega(d\Sigma) = \oint_{\sigma = \partial \Sigma} \omega(d\sigma) = F(\partial \Sigma).
+\end{equation*}
+This is the generalized Stokes' theorem and $\partial \wedge \omega$ is the exterior derivative of $\omega$. We use $\partial \wedge$ as we leave $d$ reserved for the infinitesimal regions.
 

### 2023-09-01 Notes/2023-GeneralizedEnsembleSpaces/GeneralizedEnsembleSpaces.tex
Adding notes for ensemble spaces

\documentclass[11pt]{article}

\usepackage[margin=1.25in]{geometry}
\usepackage{assumptionsofphysics}
\usepackage{hyperref}

\def\>{\rangle}
\def\<{\langle}

\DeclareMathOperator{\mix}{mix}
\DeclareMathOperator{\component}{comp}
\DeclareMathOperator{\cospan}{cospan}

\begin{document}

\title{Generalized Ensemble Spaces}

\maketitle

\begin{abstract}
The goal is to develop a generalized mathematical framework of statistical ensemble that can accommodate both the classical and quantum case, as well theories that are yet to be developed. The idea is that statistical mixtures impose a convex structure on the set of ensembles which define its geometry. This structure is tightly connected to physical entropy. All definitions must be justified on physical ground. 
\end{abstract}

\section{Convex spaces and ensembles}

\begin{remark}
	Basic definitions are taken from \url{https://ncatlab.org/nlab/show/convex+space} and \url{https://arxiv.org/abs/0903.5522}
\end{remark}

\begin{defn}
	For a real number $\lambda \in [0,1]$, $\bar{\lambda} = 1-\lambda$.
\end{defn}

\begin{defn}
	A \textbf{convex space} is a set $X$ equipped with a family of operations $c_\lambda : X \times X \to X$ where $\lambda \in [0,1]$ such that:
	\begin{itemize}
		\item $c_0(x,y)=x$
		\item $c_\lambda(x,x)=x$ for all $\lambda \in [0,1]$
		\item $c_\lambda(x,y)=c_{\bar{\lambda}}(y,x)$ for all $\lambda \in [0,1]$
		\item $c_\lambda(c_\mu(x,y), z)=c_{\lambda\mu}(x, c_{\frac{\lambda\bar{\mu}}{\overline{\lambda\mu}}}(y,z))$ for all $\lambda,\mu \in (0,1).$
	\end{itemize}
\end{defn}

\begin{remark}
	The above definition is such that, given a finite set of elements $\{x_i\}$ and a finite set of weights $\{p_i\} \in [0,1]$ such that $\sum p_i = 1$, we can define
	$$\sum p_i x_i = c_{\lambda_1}(x_1, c_{\lambda_2}(x_2, c_{\lambda_3}(x_3, \dots))$$
	and find that the sum is commutative. Such sum is called \textbf{convex combination}. The definition avoids having to define a map for each arity. However, the physical justification is more easily given in terms of the sum.

	TODO: Work out the relationship between $p_i$ and $\lambda_i$. It seems to proceed like:
	\begin{equation}
		\begin{aligned}
			\lambda_1 &= \overline{p_1} \\
			\lambda_2 &= \overline{\left(\frac{p_2}{\overline{p_1}}\right)} \\
			\lambda_2 &= \overline{\left(\frac{p_3}{\overline{p_1+p_2}}\right)}
		\end{aligned}
	\end{equation}
\end{remark}

\begin{prop}
	A convex space embeds into a real vector space with $c_\lambda(x,y) = \lambda x + \bar{\lambda}y$ if and only if
	$$ c_\lambda(x,y) = c_\lambda(x,z) \; \forall \lambda \in (0,1) \implies y = z. $$
\end{prop}
\begin{proof}
	See Theorem 4 in \url{https://arxiv.org/abs/1105.1270}.
\end{proof}

\begin{defn}
	Given a physical system, its \textbf{ensemble space} is the set of all the ensembles that characterize its state. 
\end{defn}

\begin{prop}
	Given a physical system, its \textbf{ensemble space} is a convex set. 
\end{prop}
\begin{justification}
	Given finite number of ensembles and a set of weights $\{p_i\} \in [0,1]$ such that $\sum p_i = 1$, we can in principle prepare a new ensemble that mixes the previous in the given proportions. The mixture $1x+0y$ will always give $x$, no matter the $y$. The mixture of any $x$ with itself will always return itself, no matter the weight. Mixing is commutative, so a permutation of the components corresponds to a permutation of the weights.
\end{justification}

\begin{remark}
	TODO: It is not yet clear an ensemble space must embed into a real vector space. Find a justification that proves or disproves.
\end{remark}

\begin{defn}
	Given a convex space $X$, an extreme point is an element $x \in X$ that cannot be written as convex combination of other elements of $X$.
\end{defn}

\begin{defn}
	Physically, the extreme points correspond to the \textbf{pure states}. The existence of extreme points is a major difference between convex spaces and vector spaces. In vector spaces, any element can be chose as basis for linear combination. In convex spaces, this is not the case.
\end{defn}

\subsection{Physical examples}

These are the examples that we need to keep in mind, which will need to be fully characterized in terms of convex spaces and their properties.

\subsubsection{Classical case - discrete}

The discrete classical case corresponds to a symplex (\url{https://en.wikipedia.org/wiki/Simplex}). In the finite case, the pure states $\mathcal{G} = \{g_i\}_{i=1}^{n} \subset \mathcal{E}$ are finitely many and each ensemble $e = \sum p_i g_i$ is uniquely identified by a decomposition of pure states. Effectively, each ensemble is a probability distribution over the pure states. Mathematically, each point of the space is a convex combination of the vertices. The symplex has a center point, which corresponds to the maximally mixed state, a uniform distribution over all pure states.

The countably infinite case is a bit different because not all infinite sequences $\{p_i\} \in [0,1]$ converge, and therefore cannot be normalized. As we cannot create a uniform distribution over infinitely many cases, there is no center point. Effectively, there is ``hole'' in the middle. More precisely, the space is not closed in the sense it does not contain all the limit points.

TODO: understand whether the maximally mixed state is the only limit point missing.

The uncountably infinite case is not physically relevant (i.e. no second countable discrete topology, cases are not experimentally decidable).

\subsubsection{Classical case - continuous}

In the continuous case, the space of ensembles is the space of integrable functions over a symplectic manifold (e.g.  over phase space) that integrate to one. That is, if $X$ is a symplectic manifold, then $\mathcal{E} = L^1(X, \mu)$ where $\mu=\int \omega^n$ is the Liouville measure. The pure states are the space of delta distributions $\mathcal{G} = \{\delta_x\}_{x \in X}$, but they are not in  $\mathcal{E}$: they are limit points. If we imagine a uniform distribution over a region of the symplectic manifold, we now have two limits that are not part of the ensemble space: the limit with infinite support, and the limit of zero support. In some sense, a distribution can be understood as a convex integral of the zero support limits: $\rho(x) = \int_X \rho(y) \delta_x(y) d\mu$. Even in this case, each ensemble can be understood as a probability distribution (density in this case) over all pure states.

TODO: The existence of a Liouville measure is connected to the ability to define probability densities that depend on the point, and not on the coordinate. The convex space defined finite sum. There should be an argument that shows how the existence of the Liouville measure is connected to the existence of probability density and extreme points.

TODO: The existence of the symplectic form is connected to the independence of degrees of freedom. In this context, there should be a notion of product of convex spaces, and the symplectic form should be connected to that. Find that connection.

\subsubsection{Quantum case - finite dimensional}

For a qubit, the Bloch ball is the space of ensembles $\mathcal{E}$. That is, let $\mathcal{H}$ be a two dimensional Hilbert space. The interior of the block sphere correspond to mixtures (i.e. the space of positive semi-definite Hermitian operators with trace one $M(\mathcal{H})$) while the surface corresponds to the pure states $\mathcal{G}$ (i.e. $\{ |\psi\> \<\psi| \}_{\psi \in \mathcal{H}}$). In this case, there is no single decomposition. The general finite dimensional case works similarly.

A crucial difference is that each mixed state has multiple decompositions in terms of pure states. Defining the multiple decomposition effectively defines the state space. These multiple decompositions makes the ensemble space behave in a way that is a hybrid of the classical discrete and continuous. Pure states are properly a part of the state, as in the discrete case, and we can describe each mixture in terms of finitely many states. However, the pure state are a continuum, therefore we can also define probability densities over the space, convex integrals. For example, for a single qubit, the maximally mixed state (the center of the sphere) can be equally described as the equal mixture of two opposite states (e.g. spin up and spin down, or spin left and spin right). However, it can also be described as the equal mixture of the whole sphere.

TODO: Note that complex projective are symplectic. Again, there should be an argument that the ability to define convex integrals in a coordinate invariant way requires the space to be symplectic. This would radically constrain the possible convex spaces.

\subsubsection{Quantum case - infinite dimensional}

The infinite dimensional can be both understood as the countably infinite dimensional version of the discrete case, or as the space of complex square integrable functions. As in the  (i.e. $L^2$) should work the same, except that there is no upper bound on the entropy, and therefore there is no maximally mixed state.

As before, the maximally mixed state is not the convex space.

TODO: understand the difference between Hilbert space and Schwartz space in terms of both extreme points and convex combinations/integrals.

\section{Entropy}

\begin{defn}
	Given a set of weights $\{p_i\} \in [0,1]$ such that $\sum p_i = 1$, the \textbf{entropy of the weights} is defined as $I(\{p_i\}) = - \sum p_i \log p_1 $.
\end{defn}

\begin{defn}
	Given a convex set $X$, we call \textbf{entropy} a function $S : X \to \mathbb{R}$ such that
	\begin{itemize}
		\item the entropy is strictly concave: $S(p_1\rho_1 + p_2 \rho_2) \geq p_1 S(\rho_1) + p_2 S(\rho_2)$ with the equality holding if and only if $\rho_1 = \rho_2$
		\item the entropy of a mixture is bounded by the entropy of the weights: $S(p_1\rho_1 + p_2 \rho_2) \leq I(p_1, p_2) + p_1 S(\rho_1) + p_2 S(\rho_2)$
	\end{itemize}
\end{defn}

\begin{remark}
	Note that the bounds for entropy can also be written as
	$$ 0 \leq S(p_1\rho_1 + p_2 \rho_2) - (p_1 S(\rho_1) + p_2 S(\rho_2)) \leq I(p_1, p_2).$$
	The term in the middle compares the entropy of the mixture as the average entropy of the components. It essentially tells us that the entropy can only increase during the mixture, and it can only increase up to the entropy 
\end{remark}

\subsection{Examples}

TODO: Common structure for all examples. Define what the space is, how mixing is done, and how entropy is calculated. Show that all properties are satisfied. (if we can't yet prove that $d$ is a distance function in general, at least we should prove it on each space). Show the geometrical significance of the upper bound of entropy.

\subsubsection{Classical case - discrete}

The discrete classical case corresponds to a finite dimensional symplex. The pure states $\mathcal{G} = \{g_i\}_{i=1}^{n} \subset \mathcal{E}$ are finitely many and each ensemble $e = \sum p_i g_i$ is uniquely identified by a decomposition of pure states. The entropy of each pure state is assumed to be zero.

The space for a single bit, for example, will be the line connecting the  pure states corresponding to $0$ and $1$.

\subsubsection{Classical case - continuous}

In the continuous case, the space of ensembles is the space of integrable functions over phase space (i.e. a symplectic manifold) that integrate to one. That is, if $X$ is a symplectic manifold, then $\mathcal{E} = L^1(X, \mu)$ where $\mu=\int \omega^n$ is the Liouville measure. The pure states are the space of delta distributions $\mathcal{G} = \{\delta_x\}_{x \in X}$. Note that the pure states are limit points of $\mathcal{E}$ but they are not part of it: first they are not functions and second they would have entropy minus infinity. Still, we can express an integrable function as the integral over delta distributions $\rho(x) = \int_X \rho(y) \delta_x(y) d\mu$.

For recovering this case, we will need to understand whether integrals can be defined simply as limits of sums, and whether the invariance of the measure from parametrization of the pure states can be recovered (thus recovering the notion of phase space more cleanly). A key element will be the assumption of infinitesimal reducibility, which will posit that an ensemble can always be expressed as the mixture of other ensembles.

The entropy can be used to check whether two ensembles are overlapping and define the size of the support. That is: take $\rho \in \mathcal{E}$; construct the set $\text{dis}(\rho)$ of all the ensembles that are disjoint from $\rho$; now construct the set $\text{con}(\rho)$ of all ensembles that are disjoint from all elements of $\text{dis}(\rho)$. These will be the set of ensembles that share the supports. Find the supremum of $\text{dis}(\rho)$ in terms of the entropy. If it doesn't exists, the support is infinite. If it does, the support is finite and the measure is $2^{\sup S(\text{con}(\rho))}$.

\subsubsection{Quantum case - finite dimensional}

For a qubit, the Bloch sphere is the space of ensembles $\mathcal{E}$. That is, let $\mathcal{H}$ be a two dimensional Hilbert space. The interior of the block sphere correspond to mixtures (i.e. the space of positive semi-definite Hermitian operators with trace one $M(\mathcal{H})$) while the surface corresponds to the pure states $\mathcal{G}$ (i.e. $\{ |\psi\> \<\psi| \}_{\psi \in \mathcal{H}}$). In this case, there is no single decomposition. The general finite dimensional case works similarly.

Note that the set of pure states is still a continuum, though the entropy of the space is bounded by $\log n$. Therefore we can still define integrals over pure states, though the entropy of those mixtures is still finite. A full understanding of this means understanding exactly how quantum mixes the discrete and the continuous case.

\subsubsection{Quantum case - infinite dimensional}

The infinite dimensional (i.e. $L^2$) should work the same, except that there is no upper bound on the entropy, and therefore there is no maximally mixed state.

\subsection{Convex subspaces}

\begin{defn}
	Let $X$ be a convex space and $Y \subseteq X$. We say $x \in X$ is a \textbf{mixture} of $Y$ if $x = \sum \lambda_i y_i$ for some $\{y_i\} \in Y$ and $\lambda_i \in [0,1]$ such that $\sum \lambda_i = 1$. The \textbf{mix} of $Y$, noted $\mix(Y)$, is the set of all the mixtures of $Y$.

	We say $x \in X$ is a \textbf{component} of $y \in Y$ if $\lambda x + \sum \lambda_i x_i = y$ for some $\lambda \in [0,1]$, $\{x_i\} \in X$ and $\lambda_i \in [0,1]$ such that $\sum \lambda_i = \bar{\lambda}$. The \textbf{components} of $Y$, noted $\component(Y)$, is the set of all the components of all the elements of $Y$.
\end{defn}

\begin{defn}
	Let $X$ be a convex space and $Y \subseteq X$. Then $Y$ is a \textbf{convex subspace} of $X$ if it contains all its mixtures and components. That is, $Y = \mix(Y) = \component(Y)$.
\end{defn}

\begin{defn}
	Let $X$ be a convex space and $Y \subseteq X$. The \textbf{cospan} of $Y$, noted $\cospan(Y)$ is the smallest convex subspace that contains $Y$.
\end{defn}

\begin{defn}
	Let $X$ be a convex space and $Y \subseteq X$. The \textbf{cospan} of $Y$, noted $\cospan(Y)$ is the smallest convex subspace that contains $Y$.
\end{defn}

\begin{thrm}
	(Conjecture) Let $X$ be a convex space such that, given $x_1, x_2, y_1, y_2 \in X$, $\cospan(\{x_1, x_2\}) = \cospan(\{y_1, y_2\})$. Then
	\begin{itemize}
		\item if $\cospan(x_1, x_2)$ is a segment (i.e. $B^1$), then $X$ is a symplex
		\item if $\cospan(x_1, x_2)$ is a disk (i.e. $B^2$), then $X$ is a real projective space
		\item if $\cospan(x_1, x_2)$ is a ball (i.e $B^3$), then $X$ is a complex projective space
		\item if $\cospan(x_1, x_2)$ is a 5-ball (i.e. $B^5$), then $X$ is a quaternionic projective space
	\end{itemize}
\end{thrm}

\end{document}

### 2023-09-01 Notes/2023-GeneralizedEnsembleSpaces/GeneralizedEnsembleSpaces.tex
Added more in the note for ensemble spaces

@@
 \end{remark}
+
+\begin{defn}
+	Let $X$ be a convex space.
+Let $y = \sum_i p_i x_i$ where $y, \{x_i\} \in X$ and $\lambda_i \in [0,1]$ such that $\sum \lambda_i = 1$.
+We say that $y$ is a \textbf{mixture} of $\{x_i\}$ and each $x_i$ is a \textbf{component} of $y$.
+\end{defn}
 
@@
 	The term in the middle compares the entropy of the mixture as the average entropy of the components.
-It essentially tells us that the entropy can only increase during the mixture, and it can only increase up to the entropy
+It essentially tells us that the entropy can only increase during the mixture, and it can only increase up to the entropy.
+
+	TODO: This makes the above conditions necessary; it is not clear whether they are sufficient.
+
+	Note that if an ensemble can be decomposed in multiple ways, the decomposition such that the entropy of the weights is minimal provides the least upper bound.
 \end{remark}
 
-\subsection{Examples}
+\begin{defn}
+	Two mixtures $x, y \in X$ are \textbf{disjoint} if $S(p_1\rho_1 + p_2 \rho_2) - (p_1 S(\rho_1) + p_2 S(\rho_2)) = I(p_1, p_2)$.
+\end{defn}
 
-TODO: Common structure for all examples.
-Define what the space is, how mixing is done, and how entropy is calculated.
-Show that all properties are satisfied.
-(if we can't yet prove that $d$ is a distance function in general, at least we should prove it on each space).
-Show the geometrical significance of the upper bound of entropy.
+\begin{remark}
+	Two classical distributions are disjoint if and only if their support is disjoint. Two quantum states are disjoint if and only if they are the mixture of states that belong to disjoint subspaces.
+
+	Physically, two disjoint ensembles are distinct in the sense that they will never produce overlapping results.
+\end{remark}
+
+\begin{prop}
+	The thermodynamic entropy is an entropy over an ensemble space.
+\end{prop}
+\begin{justification}
+	TODO This can be easily done in terms of information theory and statistical mechanics. An argument based on thermodynamics would be better.
+\end{justification}
+
+\subsection{Physical examples}
+
+We review the physical examples in terms of the entropy.
 
@@
 
-The discrete classical case corresponds to a finite dimensional symplex.
-The pure states $\mathcal{G} = \{g_i\}_{i=1}^{n} \subset \mathcal{E}$ are finitely many and each ensemble $e = \sum p_i g_i$ is uniquely identified by a decomposition of pure states.
-The entropy of each pure state is assumed to be zero.
+The entropy is given by $-\sum p_i \log p_i$ for both finite and countable cases.
+Therefore the entropy for each pure state is assumed to be zero.
+The entropy of the maximally mixed state is $\log n$ where $n$ is the number of pure states.
+The entropy increase as we go from pure states to the maximally mixed state.
+The level sets (i.e.
+the fibers) of the entropy form a series of concentric "shells".
 
-The space for a single bit, for example, will be the line connecting the  pure states corresponding to $0$ and $1$.
+TODO: Check that the entropy is finite for all convex combinations. Check that it diverges as we go to a limit point that is not in the ensemble space.
+
+TODO: Is it important that all pure states have zero entropy?
 
@@
 
-In the continuous case, the space of ensembles is the space of integrable functions over phase space (i.e.
-a symplectic manifold) that integrate to one.
-That is, if $X$ is a symplectic manifold, then $\mathcal{E} = L^1(X, \mu)$ where $\mu=\int \omega^n$ is the Liouville measure.
-The pure states are the space of delta distributions $\mathcal{G} = \{\delta_x\}_{x \in X}$.
-Note that the pure states are limit points of $\mathcal{E}$ but they are not part of it: first they are not functions and second they would have entropy minus infinity.
-Still, we can express an integrable function as the integral over delta distributions $\rho(x) = \int_X \rho(y) \delta_x(y) d\mu$.
-
-For recovering this case, we will need to understand whether integrals can be defined simply as limits of sums, and whether the invariance of the measure from parametrization of the pure states can be recovered (thus recovering the notion of phase space more cleanly).
-A key element will be the assumption of infinitesimal reducibility, which will posit that an ensemble can always be expressed as the mixture of other ensembles.
+The entropy is given by $- \int \rho \log \rho d\mu$ where $\mu$ is the Liouville measure and $\rho$ is the probability density over canonical coordinates.
+If a different measure is used, or if the coordinates are not canonical, the formula gives the wrong result.
 
 The entropy can be used to check whether two ensembles are overlapping and define the size of the support.
-That is: take $\rho \in \mathcal{E}$; construct the set $\text{dis}(\rho)$ of all the ensembles that are disjoint from $\rho$; now construct the set $\text{con}(\rho)$ of all ensembles that are disjoint from all elements of $\text{dis}(\rho)$.
+That is: take $\rho \in X$; construct the set $\text{dis}(\rho)$ of all the ensembles that are disjoint from $\rho$; now construct the set $\text{con}(\rho)$ of all ensembles that are disjoint from all elements of $\text{dis}(\rho)$.
 These will be the set of ensembles that share the supports.
@@
 
-\subsubsection{Quantum case - finite dimensional}
+TODO: Study the shell of zero entropy states.
+For example, it should not be path connected: all uniform distributions with support of the same finite size (in terms of Liouville measure) will have the same entropy.
+The region, however, need not be contiguous.
+Since we cannot continuously transform a single region into two disjoint regions, there will be different distributions at zero entropy that cannot be transformed continuously.
 
-For a qubit, the Bloch sphere is the space of ensembles $\mathcal{E}$.
-That is, let $\mathcal{H}$ be a two dimensional Hilbert space.
-The interior of the block sphere correspond to mixtures (i.e.
-the space of positive semi-definite Hermitian operators with trace one $M(\mathcal{H})$) while the surface corresponds to the pure states $\mathcal{G}$ (i.e.
-$\{ |\psi\> \<\psi| \}_{\psi \in \mathcal{H}}$).
-In this case, there is no single decomposition.
-The general finite dimensional case works similarly.
+\subsubsection{Quantum case}
 
-Note that the set of pure states is still a continuum, though the entropy of the space is bounded by $\log n$.
-Therefore we can still define integrals over pure states, though the entropy of those mixtures is still finite.
-A full understanding of this means understanding exactly how quantum mixes the discrete and the continuous case.
+In quantum mechanics, the entropy is the von Neumann entropy for both finite and infinite dimension.
+The entropy for the maximally mixed state is $\log n$ where $n$ is the dimensionality of the Hilbert space.
 
-\subsubsection{Quantum case - infinite dimensional}
-
-The infinite dimensional (i.e. $L^2$) should work the same, except that there is no upper bound on the entropy, and therefore there is no maximally mixed state.
+TODO: How exactly do the bounds of entropy interact with multiple decomposition?
 
@@
 
+We need to develop a theory of subspaces that map to vector subspaces in the case that the convex space is also a projective space.
+
 \begin{defn}
 	Let $X$ be a convex space and $Y \subseteq X$.
-We say $x \in X$ is a \textbf{mixture} of $Y$ if $x = \sum \lambda_i y_i$ for some $\{y_i\} \in Y$ and $\lambda_i \in [0,1]$ such that $\sum \lambda_i = 1$.
 The \textbf{mix} of $Y$, noted $\mix(Y)$, is the set of all the mixtures of $Y$.
-
-	We say $x \in X$ is a \textbf{component} of $y \in Y$ if $\lambda x + \sum \lambda_i x_i = y$ for some $\lambda \in [0,1]$, $\{x_i\} \in X$ and $\lambda_i \in [0,1]$ such that $\sum \lambda_i = \bar{\lambda}$.
 The \textbf{components} of $Y$, noted $\component(Y)$, is the set of all the components of all the elements of $Y$.
@@
 
-\begin{defn}
-	Let $X$ be a convex space and $Y \subseteq X$. The \textbf{cospan} of $Y$, noted $\cospan(Y)$ is the smallest convex subspace that contains $Y$.
-\end{defn}
+\begin{remark}
+	Note that the cospan of two elements is not simply the segment that connects them.
+That is just the mix of the two elements.
+Now we need to consider all the elements that, when mixed, can give those elements.
+If we extend the segment, the cospan will at least include those.
+
+	The cospan can even be bigger.
+Consider a ball, which is a convex space, and suppose we want to construct the cospan of a single point $A$.
+The mix of a single point is simply that point, so we have to look at the components.
+If it is on the surface, it is an extreme point, and therefore the only component is itself.
+Therefore every surface point by itself is a subspace.
+If the point is not on the surface, then take a point on the surface $B$.
+Now connect $A$ and $B$ with a line.
+This will intersect the sphere on another point $C$.
+Therefore $A$ is a convex combination of $B$ and $C$, and therefore the whole segment $BC$ is part of the cospan of $A$.
+But the same argument works with any point on the surface, therefore the cospan of any interior point of the ball is the whole ball.
+
+	TODO: The embedding of the convex space into a vector space may be related to features the cospan
+
+	TODO: The definition mimics definitions on other algebraic structure so that it can work in the infinite case. Does it?
+\end{remark}
 
@@
 
-\begin{thrm}
-	(Conjecture) Let $X$ be a convex space such that, given $x_1, x_2, y_1, y_2 \in X$, $\cospan(\{x_1, x_2\}) = \cospan(\{y_1, y_2\})$. Then
-	\begin{itemize}
-		\item if $\cospan(x_1, x_2)$ is a segment (i.e. $B^1$), then $X$ is a symplex
-		\item if $\cospan(x_1, x_2)$ is a disk (i.e. $B^2$), then $X$ is a real projective space
-		\item if $\cospan(x_1, x_2)$ is a ball (i.e $B^3$), then $X$ is a complex projective space
-		\item if $\cospan(x_1, x_2)$ is a 5-ball (i.e. $B^5$), then $X$ is a quaternionic projective space
-	\end{itemize}
-\end{thrm}
+\subsection{Physical examples}
+
+We review the physical examples in terms of subspaces.
+
+\subsubsection{Classical case - discrete}
+
+For a symplex, all subspaces will be symplexes with fewer extereme points.
+In fact, these will be the sides of simplexes with more points.
+In particular, if we take two extreme points, the cospan is the segment connecting the two.
+
+TODO We should find a characterization of the classical discrete case in terms of subspaces.
+Requiring the cospan of two extreme points to be a line is necessary but not sufficient.
+Is the converse true?
+For example, take a square.
+There are four extreme points.
+The cospan of two consecutive vertices is a segment.
+However, the cospan of two non consecutive vertices is the whole square.
+This would seem to exclude all possible flat polygons.
+
+\subsubsection{Classical case - continuous}
+
+The continuous case is trickier because the extreme points are not in the convex space.
+However, the cospan should give the set of elements that have the support in the union of the original elements.
+Note that this is related to the constuction in the previous section.
+
+\subsubsection{Quantum case}
+
+In quantum mechanics, the cospan coincides with the projective space of the subspaces of the complex vector space.
+
+TODO: Is there a relationship between entropy and cospan?
 

### 2023-09-01 Notes/2023-GeneralizedEnsembleSpaces/GeneralizedEnsembleSpaces.tex
Added mention of the JS divervence

@@
 
+\begin{remark}
+	Note that $JSD(\rho_1, \rho_2 = )S(p_1\rho_1 + p_2 \rho_2) - (p_1 S(\rho_1) + p_2 S(\rho_2))$ is the Jensen Shannon divergence (\url{https://en.wikipedia.org/wiki/Jensen%E2%80%93Shannon_divergence}).
+The square root is a distance function, and therefore makes the convex space a metric space.
+\end{remark}
+
 \subsection{Physical examples}

### 2023-09-03 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Updated version based on arxiv version comments

@@
 We suspect that at least some problems in quantum theory related to infinities may be ultimately caused by the wrong space being used.
-We strongly believe a better solution can be found, and we look at Schwartz spaces for inspiration, as, among other things, they guarantee that the expectation of all polynomials of position and momentum are finite, guarantee solution to the moment problem and are the only space closed under Fourier transform.
+We strongly believe a better solution can be found, and we look at Schwartz spaces for inspiration, as, among other things, they guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are the only space closed under Fourier transform.
 \end{abstract}
@@
 Let us first clarify the standard we believe should be used to judge the physicality of a formal definition.\cite{redei_tension_2020, north_physics_2021} Suppose we have a physical system under study in a given condition.
-A mathematical definition is \textbf{physical} if it properly characterizes a physical system.
-That is, if it can be shown that the mathematical definition is needed to capture and only capture a particular aspect of the physical system.
+A mathematical definition is \textbf{physical} if it properly characterizes the physical system.
+That is, if it can be shown that the mathematical definition is needed to capture and only capture a particular aspect of the physical system under the given condition.
 On the other hand, a mathematical definition is \textbf{unphysical} if it can be shown to require properties or operations that cannot have a physical counterpart.
@@
 Completeness forces us to include states with an infinite expectation value, which cannot be implemented physically.
-Thus the mathematical definition implicitly assumes objects that cannot be physically realized, so \textbf{the use of completeness for the space of quantum systems is unphysical.}
+Thus the mathematical definition implicitly assumes objects that cannot be physically realized, so \textbf{the use of completeness for the space of quantum systems is unphysical.}\footnote{As rigged Hilbert spaces are Hilbert spaces with additional structure, they cannot solve this problem.
+For further discussion, see the appendix.}
 
@@
 \end{itemize}
-These properties make the Schwartz space a much more reasonable candidate to capture the physics.
-The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} To again be clear, we are not proposing this as a solution.
+These properties make the Schwartz space a much more reasonable candidate to capture the physics.\cite{albert2022bosonic} The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} To again be clear, we are not proposing this as a solution.
 We are merely pointing out that alternatives do exist and should be fully explored.\footnote{Note that $C^*$-algebras model observables and therefore do not solve the problem at hand of modeling states.}
@@
 
+\section{The unphysicality of rigged Hilbert spaces}
+
+As part of the feedback we received on previous drafts, some people argued that rigged Hilbert spaces were developed to solve exactly the problems we highlight.
+They were not.
+They are designed to solve another problem, namely extending the spectral theorem to the case of continuous spectra.\cite{Madrid_2005}
+
+First of all, note that ``rigged'' in this context does not mean ``manipulated by dishonest intention'', but rather ``fully equipped and ready to go'' as, for example, a fully-rigged ship.
+Therefore, the name already tells us that rigged Hilbert spaces are Hilbert spaces with additional structure.
+They still satisfy completeness, and thus are affected by the problems presented above.
+The unphysicality of a mathematical structure cannot simply be solved by adding more structure: the unphysical elements must be removed.
+
+Rigged Hilbert spaces exist to address the fact that an operator may have eigenfunctions that are not in the Hilbert space.
+For example, the eigenstates for the operator $P = - \imath \hbar \partial_x$ are of the form $e^{\frac{xp}{-\imath\hbar}}$, and are not square integrable.
+This is due to the continuous spectrum, not the unboundedness of the operator, and is therefore an independent problem.
+We can show that with two examples.
+
+Bounded operators can have a continuous spectrum.
+For example, let $\mathcal{H}$ be the Hilbert space of all wave functions defined on the compact domain $x \in [0,1]$.
+The position operator, in this case, is bounded.
+We do not have problems of infinite expectation value.
+However, the position operator $X$ does not have eigenfunctions in $\mathcal{H}$.
+The delta Dirac distribution, in fact, is not part of the Hilbert space; a sequence of Gaussians with decreasing variance, whose limit is a delta Dirac, is not a Cauchy sequence.
+Therefore we cannot properly talk about eigenstates of $X$ in $\mathcal{H}$.
+
+Unbounded operators can have a discrete spectrum. For example, the harmonic oscillator has a discrete spectrum, given by the Hermite polynomials. All eigenfunctions are in the Hilbert space.
+
+Let us give a very rough description of the rigged Hilbert space construction.
+Given a generic Hilbert space $\mathcal{H}$, we take an operator $X$, which we assume unbounded.
+The domain $D(X)$ of the operator is a strict subset of $\mathcal{H}$.
+That is, $D(X) \subset \mathcal{H}$.
+We can then define $D(X)^*$ as the space of linear functionals of $D(X)$.\footnote{A similar construction can be done with anti-linear functionals, giving the anti-dual space.} Mathematically, $D(X)$ is the space of test functions while $D(X)^*$ is the space of tempered distributions, the dual of $D(X)$.
+This is a strict superset of $\mathcal{H}$.
+That is,
+\begin{equation}
+	D(X) \subset \mathcal{H} \subset D(X)^*.
+\end{equation}
+This is called the Gelfand triple.
+This construction allows us to talk about $\delta(x)$, as it is a tempered distribution, and therefore it can be found in $D(X)^*$.
+Note that if we chose a different operator, say $P$, we would have a different construction and a different rigged Hilbert space.
+
+The point of showing the construction is to see that the rigged Hilbert space is a Hilbert space plus two other spaces.
+Therefore, a rigged Hilbert space is physical only if all three spaces are physical.
+If the Hilbert space is not physical, neither is the rigged Hilbert space.
+
 \section{The possible physicality of Schwartz spaces}
@@
 Physically, this seems too restrictive as it excludes functions like normal distributions.
-At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
-Therefore the space of functions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}
+At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true for distributions in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
+Therefore the space of distributions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}
 
@@
 the topology of the seminorms) are enough to identify the states (i.e.
-the singletons, which are closed sets as per Hausdorff).\footnote{This is not true in classical mechanics.
-Even in the single variable case, distinct distributions can have the same moments.
-However, if we assume compact support, distinct distributions have distinct moments even in the multi-variable case.\cite{moment_problem_2017} It is curious that the moment problem in quantum mechanics works differently.} Moreover, if we have a sequence of states for which all expectations converge, we are guaranteed to converge to another state (i.e.
+the singletons, which are closed sets as per Hausdorff).\footnote{This essentially tells us that the moment problem restricted to pure states has a unique solution in quantum mechanics.
+Setting up a comparison with the classical case is delicate.
+If we restrict ourselves to classical pure states, i.e.
+points in phase space, then the solution is also unique.
+If we restrict ourselves to compact support, distinct distributions have distinct moments even in the multi-variable case.\cite{moment_problem_2017} If we restrict ourselves to integrable functions, even in the single variable case, distinct distributions can have the same moments.
+None of these comparisons seems adequate as a quantum state is a statistical object by nature, but it is not an arbitrary mixture.
+A more proper comparison would seem to restrict the problem to the space of integrable functions with zero entropy.} Moreover, if we have a sequence of states for which all expectations converge, we are guaranteed to converge to another state (i.e.
 completeness with respect to the seminorms).
@@
 First, this is not the goal of the paper, which is to show the unphysicality of Hilbert spaces.
-Second, we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The fact that the moment problems work differently in classical and quantum mechanics may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.\footnote{One of the issues is the fact that uniqueness of the moment problem works differently for classical and quantum mechanics.}
+Second, we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The lack of a clear classical parallel for the moment problem restricted to pure states may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.
 

### 2023-09-03 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Added edits from Francisco

@@
 They were not.
-They are designed to solve another problem, namely extending the spectral theorem to the case of continuous spectra.\cite{Madrid_2005}
+They are designed to solve another problem, namely extending the spectral theorem to the case of continuous spectra.\cite{Madrid_2005, reed_methods_1980, Landsman2017}
 
@@
 However, the position operator $X$ does not have eigenfunctions in $\mathcal{H}$.
-The delta Dirac distribution, in fact, is not part of the Hilbert space; a sequence of Gaussians with decreasing variance, whose limit is a delta Dirac, is not a Cauchy sequence.
+The Dirac delta distribution, in fact, is not part of the Hilbert space; a sequence of Gaussians with decreasing variance, whose limit is a Dirac delta, is not a Cauchy sequence.
 Therefore we cannot properly talk about eigenstates of $X$ in $\mathcal{H}$.
 
-Unbounded operators can have a discrete spectrum. For example, the harmonic oscillator has a discrete spectrum, given by the Hermite polynomials. All eigenfunctions are in the Hilbert space.
+Unbounded operators can have a discrete spectrum.
+For example, the harmonic oscillator has a discrete spectrum, with eigenfunctions given by the Hermite polynomials.
+All eigenfunctions are in the Hilbert space.
 
@@
 Physically, this means that we can always express a wavefunction in either the position or the momentum basis.
-Mathematically, the theory of tempered distributions, which allows us to define objects like the Dirac $\delta$-function, is built on top of the Schwartz space.
+Mathematically, the theory of tempered distributions, which allows us to define objects like the Dirac delta distribution, is built on top of the Schwartz space.
 Therefore, this space already plays a fundamental role in the definitions of the tools we already use in physics.

### 2023-10-12 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
First pass at addressing reviewers comments

@@
 Using this new approach, in this paper we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM.
-The core insight is that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.
+We start by noting that the von Neumann entropy plays a crucial role in the predictions of both quantum statistical mechanics and quantum information theory.
+We show that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.
 We find that the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all pure states must be orthogonal, which directly contradicts quantum theory.
@@
 This leads to the following expression:
-\begin{equation}
+\begin{equation}\label{prob_condition}
 	\int_\Lambda d\lambda p(k|\lambda, M) p(\lambda| P)= \textrm{tr}(\rho E_k).
@@
 \end{equation}
-All other models are \textbf{$\psi$-incomplete}.
+All other models are \textbf{$\psi$-incomplete}. For specific examples of ontological models, see \cite{Harrigan:2010}, sections 2.4.1--3.
 
@@
 
-Before proving the violation, we first introduce two results that show how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy and the von Neumann entropy.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
+Let us denote $M(\Lambda)$ the space of all probability distributions (i.e.
+all probability measures) over the space of ontic states $\Lambda$.
+Let us denote $\mathcal{H}$ the Hilbert space of the corresponding quantum system and $M(\mathcal{H})$ the space of all mixed states (i.e.
+the space of positive semi-definite trace-one Hermitian operators).
+An ontological model, then, must give us an appropriate $\rho(\lambda) \in M(\Lambda)$ for every $\rho \in M(\mathcal{H})$ such that all predictions are satisfied.
+This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$.
+Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability.\cite{aop-phys-variability} It is also the one that provides the correct link to the thermodynamic entropy.} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.
+
+We now introduce two lemmas that show how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy $H_\Lambda$ and the von Neumann entropy $H_\mathcal{H}$.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
 The second will be used to calculate the entropy on the quantum case.
@@
 \textbf{Information entropy of mixed non-overlapping probability distributions}.
-Let $\rho_1$ and $\rho_2$ be two probability density functions over a space $\Lambda$ with measure $\lambda$.\footnote{The notation of HS uses $\lambda$ as both the value and measure for integration, as often done in single variable integrals.} That is, they are the Radon-Nikodym derivatives of the respective probability measures with respect to the measure $\lambda$.
+Let $\rho_1, \rho_2 \in M(\Lambda)$ be two probability density functions over a space $\Lambda$ with measure $\lambda$.\footnote{The notation of HS uses $\lambda$ as both the value and measure for integration, as often done in single variable integrals.} That is, they are the Radon-Nikodym derivatives of the respective probability measures with respect to the measure $\lambda$.
 The entropy for each distribution is given by the usual formula\footnote{The logarithm is assumed to be in base 2.}, for example
 \begin{equation}\label{shannon_entropy}
-	H(\rho_1) = - \int_\Lambda \rho_1(\lambda) \log \rho_1(\lambda) d\lambda.
+	H_\Lambda(\rho_1) = - \int_\Lambda \rho_1(\lambda) \log \rho_1(\lambda) d\lambda.
 \end{equation}
@@
 \begin{equation}\label{entropy_nonoverlap}
-	H(\rho) = 1 + \frac{1}{2} H(\rho_1) + \frac{1}{2} H(\rho_2).
+	H_\Lambda(\rho) = 1 + \frac{1}{2} H_\Lambda(\rho_1) + \frac{1}{2} H_\Lambda(\rho_2).
 \end{equation}
@@
 	\begin{aligned}
-	H(\rho) &= H\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right) \\
+	H_\mathcal{H}(\rho) &= H_S\left(\frac{1+\sqrt{p}}{2}, \frac{1-\sqrt{p}}{2}\right) \\
 	&= - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} - \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2} ,
@@
 \end{equation}
-where the right side of the equation is the Shannon entropy calculated on the given values.
-Therefore the entropy of an equal mixture of two pure states depends only on the probability of measuring one having prepared the other.
+where the right side of the equation is the Shannon entropy calculated on the given values.\footnote{Note that $H_S$ is the ``discrete'' Shannon entropy that acts on a countable set of coefficients that sum to one.} Therefore the entropy of an equal mixture of two pure states depends only on the probability of measuring one having prepared the other.
 
-\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}
+\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}\footnote{Note that the theorem applies to all $\psi$-ontic models, including the $\psi$-supplemented ones.}
 
@@
 \begin{equation}\label{entropy_pure}
-	H(p(\lambda|P_\psi)) = H(p(\lambda|P_\phi)) = 0,
+	H_\Lambda(p(\lambda|P_\psi)) = H_\Lambda(p(\lambda|P_\phi)) = 0,
 \end{equation}
@@
 	\begin{aligned}
-		H\left(p(\lambda|\frac{1}{2}P_\psi + \frac{1}{2}P_\phi)\right) &= \\
-		H\left(\frac{1}{2}p(\lambda|P_\psi) + \frac{1}{2}p(\lambda|P_\phi)\right) 
+		H_\Lambda\left(p(\lambda|\frac{1}{2}P_\psi + \frac{1}{2}P_\phi)\right) &= \\
+		H_\Lambda\left(\frac{1}{2}p(\lambda|P_\psi) + \frac{1}{2}p(\lambda|P_\phi)\right) 
 		&= 1.
@@
 If we compare the above with eq.~\ref{entropy_mixed}, it follows that $p$ must be zero.
-In fact, the only case in which the Shannon entropy for a two-element distribution is equal to one is for the uniform distribution $H(1/2, 1/2) = 1$.
+In fact, the only case in which the Shannon entropy for a two-element distribution is equal to one is for the uniform distribution $H_S(1/2, 1/2) = 1$.
 That is, we must have that
@@
 
-First of all, information entropy is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state.
+First of all, information entropy, in all its versions, is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state.
 That is, we can check whether two states $\rho_1$ and $\rho_2$ are the same state simply by creating a mixture $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ and see whether we end up with the same entropy or not.
@@
 
-The take-home message is
+Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new\cite{Werner:2014}.
+However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics.
+For example, we can imagine preparing the direction of spin based on a classical distribution.
+Furthermore, the output of a measurement is described by classical probability.
+The temptation, then, is to simply assume that all we need to do is just put the right transition probability between the two spaces, and we are done.
+This is exactly what cannot be done and what doesn't work.
+That is
 \begin{quote}
-	\textbf{In the context of quantum mechanics, standard probability measures are allowed \emph{only} over measurement outcomes/orthogonal states.}
+	\textbf{In the context of quantum mechanics, standard probability measures are allowed over preparations and measurement outcomes, but not over states.}
 \end{quote}
-This, in fact, is exactly what we found in this paper using information theoretic considerations. In retrospect, as we said, it should not be at all surprising.
+In our approach, we exploited the fact that the entropy calculation uses the geometry of the inner product without invoking a probability of transition.
+That is, two pure states are orthogonal not because one cannot measure the first having prepared the second, but because their equal mixture raises the entropy by one bit.
+Therefore quantum mechanics is not simply defining a probability of transition between preparations and measurements: it is doing something more.
+TODO: We hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate or not in the context of quantum mechanics.
 
@@
 However, HS definitions entail that statistical mixtures are simply classical mixtures in the sense that they are captured by a standard probability measure.
-But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.
-Therefore, the HS categorization prevents one from describing quantum statistical mixing.
+But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.\footnote{The convex space formed by the mixtures is radically different\cite{bengtsson_życzkowski_2017}.
+There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics\cite{Kirkpatrick:2001,DEspagnat:2001}.} Therefore, the HS categorization prevents one from describing quantum statistical mixing.
 

### 2023-10-12 Articles/2021-HarriganSpekkens/Submissions/FoP/response1.tex
First pass at addressing reviewers comments

\documentclass[11pt, executivepaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={158mm,237mm},
 left=28mm,
 top=28mm,
 }
\usepackage{hyperref}
\hypersetup{colorlinks= true, allcolors=blue}
\setcitestyle{aysep={}}
\begin{document}

\title{\textbf{Replies to Reviewers' Comments}}

\author{The Authors}

\maketitle

\noindent We would like to thank the Editors for the opportunity to revise the manuscript; moreover, we also warmly thank both reviewers for their positive and helpful comments which led to significant changes and improvements of our paper. We appreciate that, we know that this is not an easy task.
\vspace{2mm}

In order to make the revisions immediately visible, every change in the manuscript has been written in blue so that it is highlighted with respect to the text. 
\vspace{2mm}

Finally, we hope to have properly addressed the doubts and remarks of both referees with these answers and the related changes made in the manuscript.

\section{Replies to Reviewer 1}

Reviewer 1
In the paper, the authors challenge the classification of hidden variable theories given by Harrigan and Spekkens by providing an information-theoretic argument proving that non-overlapping distributions can only represent orthogonal states. The paper is adequately written, and the essence of the argument is somewhat clear to me. However, there are some intricacies that must be addressed, which I list below.

\begin{enumerate}
	\item \textbf{Reviewer 1:} - What is the physical meaning of the entropy of the epistemic state, and why is it assumed to coincide with the entropy of quantum states? Is there an assumption I have missed (if so, my apologies)  regarding this matter? I'd appreciate if you could clarify this in your reply. Since entropy is not a measurable quantity, I'm afraid I don't see a reason for these two definitions to necessarily coincide.
	\vspace{2mm}

	\textbf{Authors:} The von Neumann entropy is linked to thermodynamic entropy and to information entropy. That is, the quantity does indeed lead to quantifiable predictions in both quantum statistical mechanics and quantum information theory. Therefore reproducing the value of the entropy is necessary to reproduce the results of quantum statistical mechanics and quantum information theory. This is mentioned a few times within the paper, though we noticed it wasn't mentioned in the introduction. We added the following sentence:

	\textcolor{blue}{We start by noting that the von Neumann entropy plays a crucial role in the predictions of both quantum statistical mechanics and quantum information theory.}

	What is interesting is that the entropy is not an observable in the standard sense: it is not a Hermitian operator and there are no eigenstates of entropy. We use the fact that the entropy of an equal mixture of a pair of pure states is uniquely defined by the probability of transition. One can go the other way: the probability of transition between two states is uniquely defined by the entropy of an equal mixture of two. So the entropy has interesting features: it leads to quantifiable predictions, it is not an observable in the standard sense and it fully captures the Born rule. That is why we focus on this quantity.

	Now, as to why we want the Shannon entropy on the epistemic state and the von Neumann entropy to coincide, we leave that to the answer of the first point of reviewer 2, as the two points are related.

\item \textbf{Reviewer 1:} - Related to the previous point, it would be pedagogical to include an example of psi-complete theory, for example, the Beltrametti-Bugajski model (or any other), to both illustrate the notion of psi-completeness, and also show how the entropic argument plays out there.
\vspace{2mm}

\textbf{Authors:} While we understand the pedagogical value of the request, we are afraid that by putting a specific example, the attention of the reader will focus on the example, and not on the general point. To avoid the issue, we have added the following sentence that refers to the original examples in the HS paper.

\textcolor{blue}{ For specific examples of ontological models, see \cite{Harrigan:2010}, sections 2.4.1--3. }

\item \textbf{Reviewer 1:} - What does your result say about psi-supplemented models?
\vspace{2mm}

\textbf{Authors:} We thank the referee for pointing out that we haven't explicitly addressed the $\psi$-supplemented case. As it is a subcase of the $\psi$-ontic one, it is automatically taken care of. We added a footnote to the theorem itself.

\textcolor{blue}{ Note that the theorem applies to all $\psi$-ontic models, including the $\psi$-supplemented ones.}

\item \textbf{Reviewer 1:} - Some of the recent criticisms (see, for example, Hance, Rarity and Ladyman) point out that the condition of non-overlapping can be interpreted as a sufficient, but not necessary for an ontic description. How would this amendment affect the results derived in the paper?
\vspace{2mm}

\textbf{Authors:} TODO

\item \textbf{Reviewer 1:} - I would also recommend to revise the Section 3 of the paper: first, outline the argument and its assumptions, and then formulate a theorem.
\vspace{2mm}

\textbf{Authors:} We revised Section 3 based on this point, the first point of reviewer 1, and the first point of reviewer 2. See the answer to the first point of reviewer 2.

\item \textbf{Reviewer 1:}  - Also, it seems to me that (9) follows from (5) rather than from the purity = zero entropy argument.
\vspace{2mm}

\textbf{Authors:} This opens up another set of problems due to the imprecision of the presentation in the original paper (and in a lot of related physics literature), where probability and probability densities are used interchangeably, though they are rather different both mathematically and physically. In general, the Shannon entropy of a uniform distribution is the logarithm of the measure on the space (this would be the $d\lambda$ used in the integral). If the space is discrete, that measure is the counting measure, and therefore we would have the logarithm of one for a distribution over a single state, which corresponds to zero entropy. However, if the space is continuous the measure is the Lebesgue measure, the Liouville measure in particular in the case of classical phase space. A single point in this case has measure zero, meaning that the entropy is minus infinity.

We specifically do not want to go down that rabbit hole, so we just adapt to the imprecise notation as best we can, while still constructing the argument in a way that we know avoids all those hidden traps. That is why we argue that pure states have zero entropy on the ground that this is what the von Neumann entropy requires.

\end{enumerate}

We hope with these answers to have properly replied to the useful comments we received from reviewer 1.

\section{Replies to Reviewer 2}

Reviewer 2
The authors present a critical analysis of the HS categorization of ontological models. When considering quantum information measures (such as the von Neumann entropy), they find a discrepancy between the entropy as computed with the ontological model and that which is directly computed with quantum theory. The conclusion is that no ontological model can reproduce quantum theory in its full strength (for example, reproducing quantum information theory). The authors go on to reconsider the PBR theorem under the light of their previous analysis.

I essentially agree with what the authors say. Also, the technical results are correct. I would just simply say: quantum theory is a strongly non-Kolmogorovian probability calculus. Therefore, all attempts to embed it into a classical probability model will give place to problems. The core of the philosophical problem is to continue the pursuit of hidden-variables. They just don’t fit naturally. The PBR theorem is nothing but an expression of this inadequacy. But, of course, there is too much confusion out there. Taking into account that confusion, the results presented become quite relevant, Sincé they help to understand better what is wrong. For these reasons, I am willing to recommend this manuscript for publication.

But I have come across comments regarding the author’s arguments. I don’t think that the manuscript is ready to recommend acceptance. I think that, if the authors take care of the observations listed below, the manuscript results will become stronger.

\begin{enumerate}
	\item \textbf{Reviewer 2:} - (1)    Main problem: the functions H appearing in (7) and (8) are not the same. One of them -let us call it $H_S$, “Shannon’s Entropy”- has as domain the set of classical probability densities. The second one, let us call it $H_vn$, ha sas domain the set of density operators representing quantum states. A defender of the HS formalism might say: “Why should these functions have the same outputs?”. Imposing that they should have the same output is a very natural condition that any reasonable physicists would follow. A quantum state will give place to a classical probability in the Lambda space. Thus, even if the domains are different, the outputs should be the same. But of course, this leads to problems, as the authors clearly explain. My proposal is to mention that they are not the same function, to use different names for these functions, and to remark that it is natural for a physicist to demand that they should yield the same result.  
	\vspace{2mm}

	\textbf{Authors:} Both this comment and the previous comments from the other referee tells us that the motivation and setting for the theorem can be greatly improved. We have added an introductory paragraph where the distinction between the functions and the overall motivation to require their accordance is better highlighted. We hope that this would address potential objections as reviewer 2 notes.

	\textcolor{blue}{Let us denote $M(\Lambda)$ the space of all probability distributions (i.e. all probability measures) over the space of ontic states $\Lambda$. Let us denote $\mathcal{H}$ the Hilbert space of the corresponding quantum system and $M(\mathcal{H})$ the space of all mixed states (i.e. the space of positive semi-definite trace-one Hermitian operators). An ontological model, then, must give us an appropriate $\rho(\lambda) \in M(\Lambda)$ for every $\rho \in M(\mathcal{H})$ such that all predictions are satisfied. This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$. Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability.\cite{aop-phys-variability} It is also the one that provides the correct link to the thermodynamic entropy.} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.
	}

	We also made other small modification and adjustments to the notation in the rest of the sections.

\item \textbf{Reviewer 2:} -   (2)    On page 4, there is a take-home message. Why not use the more traditional take-home message, which is: “Quantum theory is not a classical probability calculus”. There is no novelty in the take-home message. By the way, some citations to previous works could be included. It comes to my mind the discussion between Tim Maudlin and R. Werner in JPA, regarding the correct interpretation of Bell’s theorem. Werner clearly says that the quantum state space is not a simplex (while the classical probability space is a simplex).
\vspace{2mm}

\textbf{Authors:} The issue is that just saying “Quantum theory is not a classical probability calculus” does not seem to home in on the problem, as people seem to still be confused. Admittedly, we do use a fair amount of classical probability concepts in quantum mechanics as well, and a lot of calculations are done in similar fashion. So we are trying to home in more precisely on what is and what is not allowed, though clearly what we wrote does not reflect that.

As we see it, the issue is that we do use standard classical probability when describing preparations (i.e. the settings of the preparation device), and the machinery of quantum mechanics does spit out a classical probability for measurements. So, we believe that we can understand those as classical. Quantum states sit in between, and it's those that cannot support classical probability calculus.

Based on the suggestion, we rewrote the take-home-message paragraph as follows, including the citation to Werner's work.

\item \textbf{Reviewer 2:} - (3)    In page 5, in the conclusions, the authors mention that “quantum mixtures are non-classical”. But again, this is very well known. I think that the distinction between proper vs improper mixtures should be mentioned here. There is a nice discussion between Kirkpatrick and D’Espagnat on the arxives about this.
\vspace{2mm}

\textbf{Authors:} We do agree that the fact that the space of classical and quantum mixtures are different is very well known. We added a footnote to provide some references, including to the discussion on the distinguishability of proper and improper mixtures.

\textcolor{blue} {The convex space formed by the mixtures is radically different\cite{bengtsson_życzkowski_2017}. There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics\cite{Kirkpatrick:2001,DEspagnat:2001}.
}

\item \textbf{Reviewer 2:} - (4)    I really think that a discussion about the interpretations of quantum theory should be included here. I’ve been involved in discussions with Bohmians and Everettians regarding the PBR theorem. They are happy with it, because they need to claim that the wave function is somehow real. When I say that the hypothesis of the PBR theorem uses classical probability (and they shouldn’t), they just say that classical probability governing the Lambdas is a very natural requirement. I think that this manuscript could be helpful for young people to understand why it is so artificial to try to inject ontological models in the quantum domain.
\vspace{2mm}

\textbf{Authors:} TODO.

\item \textbf{Reviewer 2:} - (5)    Last but no least. There is usually a very rough reasoning when the dichotomy psi-ontic vs psi-epistemic is posed. The discussion is much more subtle. There is a very nice article by Hans Halvorson posing the discussion in a more sophisticated way (To Be a Realist about Quantum Theory).
\vspace{2mm}

\textbf{Authors:} TODO.

\end{enumerate}

\noindent With these answers we hope to have properly responded to the comments of reviewer 2, which have been very useful to improve the quality of the present essay.

\clearpage
\bibliographystyle{apalike}
\bibliography{../bibliography}
\end{document}

### 2023-10-13 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
Other minor response edits

@@
 Therefore quantum mechanics is not simply defining a probability of transition between preparations and measurements: it is doing something more.
-TODO: We hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate or not in the context of quantum mechanics.
+While it is not in the scope of this article to articulate precisely what this ``more'' is, which will be the focus of future work, we hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate and not in the context of quantum mechanics.
 

### 2023-10-13 Articles/2021-HarriganSpekkens/Submissions/FoP/response1.tex
Other minor response edits

@@
 
-\noindent We would like to thank the Editors for the opportunity to revise the manuscript; moreover, we also warmly thank both reviewers for their positive and helpful comments which led to significant changes and improvements of our paper.
-We appreciate that, we know that this is not an easy task.
+\noindent We are grateful to the Editors and both reviewers for the constructive review process on a subject that seems to attract a lot of animosity.
+In particular, we would like to thank the reviewers for their the positive and helpful comments, which led to meaningful changes and improvements of our paper.
 \vspace{2mm}
@@
 
-Finally, we hope to have properly addressed the doubts and remarks of both referees with these answers and the related changes made in the manuscript.
+We hope to have addressed the remarks of both referees with these answers and the related changes made in the manuscript.
 
 \section{Replies to Reviewer 1}
-
-Reviewer 1
-In the paper, the authors challenge the classification of hidden variable theories given by Harrigan and Spekkens by providing an information-theoretic argument proving that non-overlapping distributions can only represent orthogonal states.
-The paper is adequately written, and the essence of the argument is somewhat clear to me.
-However, there are some intricacies that must be addressed, which I list below.
 
@@
 
-Reviewer 2
-The authors present a critical analysis of the HS categorization of ontological models.
-When considering quantum information measures (such as the von Neumann entropy), they find a discrepancy between the entropy as computed with the ontological model and that which is directly computed with quantum theory.
-The conclusion is that no ontological model can reproduce quantum theory in its full strength (for example, reproducing quantum information theory).
-The authors go on to reconsider the PBR theorem under the light of their previous analysis.
-
-I essentially agree with what the authors say.
-Also, the technical results are correct.
+\begin{enumerate}
+	\item \textbf{Reviewer 2:} Also, the technical results are correct.
 I would just simply say: quantum theory is a strongly non-Kolmogorovian probability calculus.
@@
 For these reasons, I am willing to recommend this manuscript for publication.
+	\vspace{2mm}
 
-But I have come across comments regarding the author’s arguments.
-I don’t think that the manuscript is ready to recommend acceptance.
-I think that, if the authors take care of the observations listed below, the manuscript results will become stronger.
+	\textbf{Authors:} We concur with the assessment.
+We essentially constructed an alternative argument to outline the failure from a different direction.
+We do agree that, in the end, it is exactly the same failure of classical probability calculus, and we do mention it.
+The reason that we don't simply say that is that those that already disregard that reasoning would simply disregard our result, because it would argue a point they are already disregarding.
 
-\begin{enumerate}
 	\item \textbf{Reviewer 2:} - (1)    Main problem: the functions H appearing in (7) and (8) are not the same.
@@
 
+\textcolor{blue} {
+	Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new[33].
+However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics.
+For example, we can imagine preparing the direction of spin based on a classical distribution.
+Furthermore, the output of a measurement is described by classical probability.
+The temptation, then, is to simply assume that all we need to do is just put the right transition probability between the two spaces, and we are done.
+This is exactly what cannot be done and what doesn't work.
+That is
+	\begin{quote}
+		\textbf{In the context of quantum mechanics, standard probability measures are allowed over preparations and measurement outcomes, but not over states.}
+	\end{quote}
+	In our approach, we exploited the fact that the entropy calculation uses the geometry of the inner product without invoking a probability of transition.
+That is, two pure states are orthogonal not because one cannot measure the first having prepared the second, but because their equal mixture raises the entropy by one bit.
+Therefore quantum mechanics is not simply defining a probability of transition between preparations and measurements: it is doing something more.
+While it is not in the scope of this article to articulate precisely what this ``more'' is, which will be the focus of future work, we hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate and not in the context of quantum mechanics.
+}
+
 \item \textbf{Reviewer 2:} - (3)    In page 5, in the conclusions, the authors mention that “quantum mixtures are non-classical”.
@@
 
-\textcolor{blue} {The convex space formed by the mixtures is radically different\cite{bengtsson_życzkowski_2017}.
-There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics\cite{Kirkpatrick:2001,DEspagnat:2001}.
+\textcolor{blue} {The convex space formed by the mixtures is radically different[34].
+There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics[35-36].
 }

### 2023-10-14 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
Changes from Andrea

--- 2023-10-12 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
@@
 \documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
-%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
-%\documentclass[10pt,twocolumn,letterpaper]{article}
-
 \usepackage{amsmath}
@@
 
-In 2010 Harrigan and Spekkens (HS) proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}.
+In 2010 Harrigan and Spekkens (HS) proposed a \textcolor{magenta}{formal} classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}.
 While their original aim was to clarify Einstein's view of Quantum Mechanics (QM), the HS framework has been widely employed in the literature not only to categorize different formulations of QM, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Ladyman:2021} for critical discussions).
@@
 
-Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information.
-On the other, scholars recently showed that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2021, DeBrota:2019}).
+Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, \textcolor{magenta}{some authors believe that it rules out interpretations of QM where $\psi$ merely represents information.
+On the other hand, it has been recently showed by other scholars that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2020b, Oldofredi:2021, DeBrota:2019}).}
 
 While the discussion has been focused mainly on whether $\psi$-epistemic models are problematic or whether the PBR argument itself is problematic, we ask a different question: what if the underlying HS classification itself is problematic?
-In particular, the framework assumes that once pure states are fully characterized, mixed states are constructed simply as a standard (i.e.
-classical) statistical mixture.
+In particular, the framework assumes that once pure states are fully characterized, mixed states are constructed simply as a standard (i.e.\ classical) statistical mixture.
 However, as it is well-known, classical and quantum mixtures have different properties.
@@
 
-Using this new approach, in this paper we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM.
-We start by noting that the von Neumann entropy plays a crucial role in the predictions of both quantum statistical mechanics and quantum information theory.
-We show that the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.
+\textcolor{magenta}{Employing} this new approach, in this paper we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM.
+\textcolor{blue}{We start by noting that the von Neumann entropy plays a crucial role in the predictions of both quantum statistical mechanics and quantum information theory.
+We show that} the degree of overlap between two probability distributions affects how the entropy of their mixture relates to the entropy of the components.
 We find that the lack of overlap of epistemic states---a necessary condition for a model to be $\psi$-ontic---requires that all pure states must be orthogonal, which directly contradicts quantum theory.
@@
 \end{equation}
-All other models are \textbf{$\psi$-incomplete}. For specific examples of ontological models, see \cite{Harrigan:2010}, sections 2.4.1--3.
+All other models are \textbf{$\psi$-incomplete}. \textcolor{blue}{For specific examples of ontological models, see \cite{Harrigan:2010}, sections 2.4.1--3.}
 
@@
 Temperature is a property of the preparation and therefore an \emph{epistemic property}.
+
+\textcolor{magenta}{After having introduced the relevant definitions of the HS ontological model framework, it should be noted that it has been subject to several criticisms, as already said in the previous section.
+For instance, Oldofredi \& Lopez in \cite{Oldofredi:2020b} analyzed the implicit assumptions made by Harrigan \& Spekkens in order to define their categorization, and showed that relevant interpretations of quantum theory---such as the statistical, relational and perspectival readings of the quantum formalism---cannot be classified according to the HS approach, given that their principles are in tension with those employed in \cite{Harrigan:2010}.
+Similarly, Hance, Rarity and Ladyman argued in \cite{Ladyman:2021} that the dichotomy ``ontic vs.\ epistemic'' proposed by HS is artificial, since there are cases in which the quantum state can represent both ontic properties of quantum systems as well as agents' knowledge of them.
+Thus, they claim, the definitions of $\psi$-ontic and $\psi$-epistemic introduced a few lines above do not capture the intuitive ideas with which physicists and philosophers alike use the terms ``ontic'' and ``epistemic''.}
+
+\textcolor{magenta}{While we agree with the main messages of these critical responses to the HS framework, the argument offered in this essay follows a different and complementary strategy.
+In fact, instead of criticizing the assumptions used by Harrigan \& Spekkens, we consider them valid (for the sake of the argument) and derive a no-go theorem for $\psi$-ontic models directly from them, as we are going to show in the next section.}
 
@@
 
-Let us denote $M(\Lambda)$ the space of all probability distributions (i.e.
+\textcolor{blue}{Let us denote $M(\Lambda)$ the space of all probability distributions (i.e.
 all probability measures) over the space of ontic states $\Lambda$.
@@
 This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$.
-Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability.\cite{aop-phys-variability} It is also the one that provides the correct link to the thermodynamic entropy.} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.
+Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{\textcolor{blue}{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability \cite{Carcassi:2021}.
+It is also the one that provides the correct link to the thermodynamic entropy.}} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.}
 
-We now introduce two lemmas that show how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy $H_\Lambda$ and the von Neumann entropy $H_\mathcal{H}$.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
+\textcolor{blue}{We now introduce two lemmas that} show how the entropy behaves when mixing two states, both for the Shannon/Gibbs entropy $H_\Lambda$ and the von Neumann entropy $H_\mathcal{H}$.\footnote{Full calculations with all steps are included in the appendix.} The first will be used to calculate entropy of statistical mixtures of epistemic states, since equation \ref{epistemic_mixing} tells us the mixture is done according to a standard (classical) measure of probability.
 The second will be used to calculate the entropy on the quantum case.
@@
 \end{equation}
-where the right side of the equation is the Shannon entropy calculated on the given values.\footnote{Note that $H_S$ is the ``discrete'' Shannon entropy that acts on a countable set of coefficients that sum to one.} Therefore the entropy of an equal mixture of two pure states depends only on the probability of measuring one having prepared the other.
+where the right side of the equation is the Shannon entropy calculated on the given values.\footnote{\textcolor{blue}{Note that $H_S$ is the ``discrete'' Shannon entropy that acts on a countable set of coefficients that sum to one.}} Therefore the entropy of an equal mixture of two pure states depends only on the probability of measuring one having prepared the other.
 
-\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}\footnote{Note that the theorem applies to all $\psi$-ontic models, including the $\psi$-supplemented ones.}
+\textbf{Theorem: Since non-overlapping distributions can only represent orthogonal states, $\psi$-ontic models cannot be consistent with quantum theory.}\footnote{\textcolor{blue}{Note that the theorem applies to all $\psi$-ontic models, including the $\psi$-supplemented ones.}}
 
@@
 
-First of all, information entropy, in all its versions, is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state.
+First of all, information entropy, \textcolor{blue}{in all its versions,} is strictly concave, meaning that $H(p \rho_1 + (1-p) \rho_2) \geq p H(\rho_1) + (1-p) H(\rho_2)$ with the equality valid only if $\rho_1$ and $\rho_2$ are the same exact state.
 That is, we can check whether two states $\rho_1$ and $\rho_2$ are the same state simply by creating a mixture $\rho = \frac{1}{2} \rho_1 + \frac{1}{2} \rho_2$ and see whether we end up with the same entropy or not.
@@
 In our result, we saw that the use of standard measure theory when creating mixtures as in equation \ref{epistemic_mixing} generates a problem because the convex space of classical distributions is significantly different from the convex space of quantum mixed states.
-For this case, ontological models are ``not quantum enough.''
+For this case, ontological models are ``not quantum enough''.
 
@@
 This assumes that a composite system is describable using a standard classical probability measure, which, in turn, allows us to express the joint probability distribution of observables of different subsystems.
-This, in general, will fail because it is ``not quantum enough.''
+\textcolor{magenta}{Again, this in general will fail because it is ``not quantum enough''.}
 
@@
 
-Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new\cite{Werner:2014}.
+\textcolor{blue}{Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new \cite{Werner:2014, Pitowsky:1989}.
 However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics.
@@
 The temptation, then, is to simply assume that all we need to do is just put the right transition probability between the two spaces, and we are done.
-This is exactly what cannot be done and what doesn't work.
-That is
+This is exactly what cannot be done and what does not work.
+That is}
 \begin{quote}
-	\textbf{In the context of quantum mechanics, standard probability measures are allowed over preparations and measurement outcomes, but not over states.}
+	\textcolor{blue}{\textbf{In the context of quantum mechanics, standard probability measures are allowed over preparations and measurement outcomes, but not over states.}}
 \end{quote}
-In our approach, we exploited the fact that the entropy calculation uses the geometry of the inner product without invoking a probability of transition.
+\textcolor{blue}{In our approach, we exploited the fact that the entropy calculation uses the geometry of the inner product without invoking a probability of transition.
 That is, two pure states are orthogonal not because one cannot measure the first having prepared the second, but because their equal mixture raises the entropy by one bit.
 Therefore quantum mechanics is not simply defining a probability of transition between preparations and measurements: it is doing something more.
-TODO: We hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate or not in the context of quantum mechanics.
+While it is not in the scope of this article to articulate precisely what this ``more'' is, which will be the focus of future work, we hope that this additional insight may provide more precise guidance as to where classical probability calculus is appropriate and not in the context of quantum mechanics.}
 
@@
 
-The key problem of the HS categorization is that it is ``not quantum enough.'' The framework tries to hide the non-classicality at the level of each ontic state, which, by construction, cannot be decomposed further using a standard probability measure.
+The key problem of the HS categorization is that it is ``not quantum enough''.
+The framework in fact tries to hide the non-classicality at the level of each ontic state which, by construction, cannot be decomposed further using a standard probability measure.
 However, HS definitions entail that statistical mixtures are simply classical mixtures in the sense that they are captured by a standard probability measure.
-But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.\footnote{The convex space formed by the mixtures is radically different\cite{bengtsson_życzkowski_2017}.
-There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics\cite{Kirkpatrick:2001,DEspagnat:2001}.} Therefore, the HS categorization prevents one from describing quantum statistical mixing.
+But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.\footnote{\textcolor{blue}{The convex space formed by the mixtures is radically different \cite{Bengtsson:2017}.
+There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics \cite{Kirkpatrick:2001}, \cite{DEspagnat:2001}.}} Therefore, the HS categorization prevents one from describing quantum statistical mixing.
 
-Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are untenable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions.
+Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are untenable not because there is a problem with epistemic models in particular, but rather there is an underlying problem in the HS definitions.
 Specifically, the problem of the HS categorization is the use of standard probability theory in composite systems.
@@
 
-As a final note, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models.
-To our knowledge, however, none of them is able to reproduce all results from quantum statistical mechanics (e.g.
-derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g.
-derive the Holevo bound).
+\textcolor{magenta}{Referring to this, two final remarks are in order: on the one hand, our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory.
+Indeed, many scholars use the PBR result to vindicate the empirical adequacy of a given framework---especially those in which the quantum state is somehow supposed to represent real physical objects---or to rule out a specific reading of the quantum formalism.
+Here we warn physicists and philosophers that no argument resting on Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.}
+
+\textcolor{magenta}{To this regard, however, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models.
+To our knowledge, however, none of them is fully empirically equivalent to quantum theory, i.e.\ is able to reproduce all results from quantum statistical mechanics (e.g., derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g., derive the Holevo bound).
 Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.
+In sum, we not only agree with Hance, Rarity and Ladyman in saying that HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:}
+\begin{quote}
+\textcolor{magenta}{\textbf{If a certain interpretation of the quantum formalism\footnote{\textcolor{magenta}{As for instance de Broglie-Bohm theory, Everett's relative state formulation, Relational quantum mechanics or the Many-Worlds perspective \emph{etc.}.}} is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}}
+\end{quote}
+\noindent \textcolor{magenta}{In turn, the logical conjunction of our result and the PBR theorem entails that no interpretation of quantum theory can be correctly categorized according to the HS approach.}
+
+\textcolor{magenta}{On the other hand, and we conclude, another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned.
+In their essay Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ is put into a  one-to-one correspondence with the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system.
+This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models.
+However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in literature does entail several conundrums that should be avoided---cf.\  \cite{Halvorson:2019} for details.
+To this regard, in fact, HS do not clearly spelled out what they mean when it is asserted that $\psi$ ``represents'' a real property of physical systems.
+Alternatively, to claim that $\psi$-ontic models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered: for instance it is not specified in which way $\psi$ represents the ontic state, is it a direct or indirect representation?
+If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$?
+To these and other issues Harrigan \& Spekkens do not reply.}
+
+\textcolor{magenta}{Therefore, not only their categorization is shown to be empty, but also it does not offer a rigorous philosophical ground where to cast the interpretational debate.}
 

### 2023-10-14 Articles/2021-HarriganSpekkens/Submissions/FoP/response1.tex
Changes from Andrea

@@
 
-\textcolor{blue}{ For specific examples of ontological models, see \cite{Harrigan:2010}, sections 2.4.1--3. }
+\textcolor{blue}{ For specific examples of ontological models, see [1], sections 2.4.1--3. }
 
@@
 
-\textbf{Authors:} TODO
+\textbf{Authors:} The referee highlights an interesting point in this comment.
+In the mentioned work Hance, Rarity and Ladyman criticize the assumptions on which Harrigan \& Spekkens classification rests on, and show that the dichotomy between $\psi$-ontic and $\psi$-epistemic models is artificial.
+In the revised manuscript we added a new paragraph at the end of section II mentioning this paper and how our argumentative strategy is complementary w.r.t.\ that employed by Hance, Rarity and Ladyman:
+
+\textcolor{magenta}{After having introduced the relevant definitions of the HS ontological model framework, it should be noted that it has been subject to several criticisms, as already said in the previous section.
+For instance, Oldofredi \& Lopez in [10] analyzed the implicit assumptions made by Harrigan \& Spekkens in order to define their categorization, and showed that relevant interpretations of quantum theory---such as the statistical, relational and perspectival readings of the quantum formalism---cannot be classified according to the HS approach, given that their principles are in tension with those employed in [1].
+Similarly, Hance, Rarity and Ladyman argued in [11] that the dichotomy ``ontic vs.\ epistemic'' proposed by HS is artificial, since there are cases in which the quantum state can represent both ontic properties of quantum systems as well as agents' knowledge of them.
+Thus, they claim, the definitions of $\psi$-ontic and $\psi$-epistemic introduced a few lines above do not capture the intuitive ideas with which physicists and philosophers alike use the terms ``ontic'' and ``epistemic''.}
+
+\textcolor{magenta}{While we agree with the main messages of these critical responses to the HS framework, the argument offered in this essay follows a different and complementary strategy.
+In fact, instead of criticizing the assumptions used by Harrigan \& Spekkens, we consider them valid (for the sake of the argument) and derive a no-go theorem for $\psi$-ontic models directly from them, as we are going to show in the next section.}
+
+The result obtained by Hance, Rarity and Ladyman is not in contrast with ours since both essays aim at showing the inadequacy of the HS in order to correctly classify interpretations of quantum mechanics.
+However, while it can be true that  the ``condition of non-overlapping can be interpreted as a sufficient, but not necessary for an ontic description'' as the referee correctly says, in our manuscript we consider the HS definitions as true in order to derive a no-go theorem for $\psi$-ontic directly from them.
 
@@
 But, of course, there is too much confusion out there.
-Taking into account that confusion, the results presented become quite relevant, Sincé they help to understand better what is wrong.
+Taking into account that confusion, the results presented become quite relevant, Since they help to understand better what is wrong.
 For these reasons, I am willing to recommend this manuscript for publication.
@@
 This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$.
-Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability.\cite{aop-phys-variability} It is also the one that provides the correct link to the thermodynamic entropy.} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.
-	}
+Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.
+[Footnote: Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability [30].
+It is also the one that provides the correct link to the thermodynamic entropy.] Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.}
 
@@
 \textcolor{blue} {
-	Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new[33].
+	Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new [33].
 However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics.
@@
 I think that the distinction between proper vs improper mixtures should be mentioned here.
-There is a nice discussion between Kirkpatrick and D’Espagnat on the arxives about this.
+There is a nice discussion between Kirkpatrick and D’Espagnat on the arXiv about this.
 \vspace{2mm}
@@
 
-\textcolor{blue} {The convex space formed by the mixtures is radically different[34].
-There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics[35-36].
+\textcolor{blue} {The convex space formed by the mixtures is radically different [34].
+There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics [35-36].
 }
@@
 
-\textbf{Authors:} TODO.
+\textbf{Authors:} The reviewer raises a very important point that was not adequately discussed in the original version of the manuscript.
+In the revised version of the paper we added a new paragraph in the conclusion section saying that the HS framework should not be employed in order to assess the viability of interpretations of the quantum formalism---neither to claim that a certain reading of QM is empirically adequate, not to rule out a specific view:
+
+\textcolor{magenta}{Referring to this, two final remarks are in order: on the one hand, our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory.
+Indeed, many scholars use the PBR result to vindicate the empirical adequacy of a given framework---especially those in which the quantum state is somehow supposed to represent real physical objects---or to rule out a specific reading of the quantum formalism.
+Here we warn physicists and philosophers that no argument resting on Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.}
+
+\textcolor{magenta}{To this regard, however, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models.
+To our knowledge, however, none of them is fully empirically equivalent to quantum theory, i.e.\ is able to reproduce all results from quantum statistical mechanics (e.g., derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g., derive the Holevo bound).
+Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.
+In sum, we not only agree with Hance, Rarity and Ladyman in saying that HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:}
+\begin{quote}
+\textcolor{magenta}{\textbf{If a certain interpretation of the quantum formalism is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}}
+\end{quote}
+\noindent \textcolor{magenta}{In turn, the logical conjunction of our result and the PBR theorem entails that no interpretation of quantum theory can be correctly categorized according to the HS approach.}
+
+Another general problem that we see---similar to that mentioned by referee 2---is that the notions of ``ontic'' and ``epistemic'' have always been liberally employed when interpreting quantum theories, but now it is commonly believed that HS provide a set of formal definitions capturing the meaning of such terms.
+In this way, many authors think to substantiate their views with the aid of HS classification.
+We believe that it will be very hard to deconstruct such an habit, however, our essay can be another step to achieve this goal.
+We explicitly discussed a paper of Hance, Rarity and Ladyman in section II, where this issue is openly addressed.
+For more details cf.\ the response to the fourth comment of referee 1.
 
@@
 
-\textbf{Authors:} TODO.
+\textbf{Authors:} Again, we agree with the referee's comment.
+We firmly believe that the philosophical background of HS is problematic, as one of us argued in [10].
+Following the suggestion we added another final paragraph to the conclusion section highlighting that the realist view about the quantum state---its ``onticity''---contained in $\psi$-ontic models is poorly defined:
 
+\textcolor{magenta}{On the other hand, and we conclude, another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned.
+In their essay Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ is put into a  one-to-one correspondence with the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system.
+This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models.
+However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in literature does entail several conundrums that should be avoided---cf.\  \cite{Halvorson:2019} for details.
+To this regard, in fact, HS do not clearly spelled out what they mean when it is asserted that $\psi$ ``represents'' a real property of physical systems.
+Alternatively, to claim that $\psi$-ontic models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered: for instance it is not specified in which way $\psi$ represents the ontic state, is it a direct or indirect representation?
+If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$?
+To these and other issues Harrigan \& Spekkens do not reply.}
+
+\noindent We did not included a more elaborated discussion of the latter point since the main focus of the paper is to prove the technical statement that $\psi$-ontic models cannot fully reproduce quantum theory.
+However, we will address the philosophical implications of our analysis in another philosophically oriented paper that is in preparation.
 \end{enumerate}

### 2023-10-15 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
Reviewed Andrea's changes with Christine

@@
 
-Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, \textcolor{magenta}{some authors believe that it rules out interpretations of QM where $\psi$ merely represents information.
-On the other hand, it has been recently showed by other scholars that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2020b, Oldofredi:2021, DeBrota:2019}).}
+Such a theorem had a remarkable resonance \cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}, and questions about its actual meaning are still discussed today: on the one hand, \textcolor{magenta}{some authors believe that it rules out interpretations of QM where $\psi$ merely represents information.
+On the other hand, it has recently been shown by other scholars that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument \cite{Ben:2017, Rizzi:2018, Oldofredi:2020b, Oldofredi:2021, DeBrota:2019}.}
 
@@
 
-\textcolor{magenta}{While we agree with the main messages of these critical responses to the HS framework, the argument offered in this essay follows a different and complementary strategy.
+\textcolor{magenta}{While we agree with the main messages of these critical responses to the HS framework, the argument offered in this paper follows a different and complementary strategy.
 In fact, instead of criticizing the assumptions used by Harrigan \& Spekkens, we consider them valid (for the sake of the argument) and derive a no-go theorem for $\psi$-ontic models directly from them, as we are going to show in the next section.}
@@
 
+\textcolor{blue}{\section{On interpretations}}
+
+\textcolor{magenta}{Our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory.
+Indeed, many scholars use the PBR result to vindicate the empirical adequacy of a given framework---especially those in which the quantum state is somehow supposed to represent real physical objects---or to rule out a specific reading of the quantum formalism.
+Here we warn physicists and philosophers that no argument resting on the Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.}
+
+\textcolor{magenta}{Therefore, we not only agree with Hance, Rarity and Ladyman in saying that the HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:}
+\begin{quote}
+	\textcolor{magenta}{\textbf{If a certain interpretation of the quantum formalism\footnote{\textcolor{magenta}{As for instance de Broglie-Bohm theory, Everett's relative state formulation, relational quantum mechanics or the many-worlds perspective \emph{etc.}.}} is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}}
+\end{quote}
+\noindent \textcolor{magenta}{In turn, the logical conjunction of our result and the PBR theorem entails that no interpretation of quantum theory can be correctly categorized according to the HS approach.}
+
+\textcolor{magenta}{Another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned.
+In their paper Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ fully identifies the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system.
+This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models.
+However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in the literature entails several conundrums that should be avoided---cf.\  \cite{Halvorson:2019} for details.
+With regard to this, the HS framework does not clearly spell out what is meant by the fact that $\psi$ ``represents'' a real property of physical systems.
+In other words, to claim that in $\psi$-complete models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered.
+For instance, it is not specified in which way $\psi$ represents the ontic state: is it a direct or indirect representation?
+If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$?
+On these and other issues the Harrigan \& Spekkens framework does not give guidance.}
+
+\textcolor{magenta}{Therefore, not only is their categorization shown to be empty, but also it does not offer a rigorous philosophical ground on which to cast the interpretational debate.}
+
 \section{Conclusion}
@@
 But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.\footnote{\textcolor{blue}{The convex space formed by the mixtures is radically different \cite{Bengtsson:2017}.
-There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics \cite{Kirkpatrick:2001}, \cite{DEspagnat:2001}.}} Therefore, the HS categorization prevents one from describing quantum statistical mixing.
+There is also debate as to whether proper and improper mixtures are distinguishable in the context of quantum mechanics \cite{Kirkpatrick:2001}, \cite{DEspagnat:2001}.}} Therefore, the HS categorization prevents one from describing quantum statistical mixing.
 
@@
 If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories, an argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.
+\textcolor{magenta}{The direct consequence of this is that the HS categorization should not be employed as a tool to classify quantum interpretations or to draw philosophically sound conclusions on the nature of quantum mechanics.}
 
-\textcolor{magenta}{Referring to this, two final remarks are in order: on the one hand, our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory.
-Indeed, many scholars use the PBR result to vindicate the empirical adequacy of a given framework---especially those in which the quantum state is somehow supposed to represent real physical objects---or to rule out a specific reading of the quantum formalism.
-Here we warn physicists and philosophers that no argument resting on Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.}
-
-\textcolor{magenta}{To this regard, however, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models.
-To our knowledge, however, none of them is fully empirically equivalent to quantum theory, i.e.\ is able to reproduce all results from quantum statistical mechanics (e.g., derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g., derive the Holevo bound).
-Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.
-In sum, we not only agree with Hance, Rarity and Ladyman in saying that HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:}
-\begin{quote}
-\textcolor{magenta}{\textbf{If a certain interpretation of the quantum formalism\footnote{\textcolor{magenta}{As for instance de Broglie-Bohm theory, Everett's relative state formulation, Relational quantum mechanics or the Many-Worlds perspective \emph{etc.}.}} is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}}
-\end{quote}
-\noindent \textcolor{magenta}{In turn, the logical conjunction of our result and the PBR theorem entails that no interpretation of quantum theory can be correctly categorized according to the HS approach.}
-
-\textcolor{magenta}{On the other hand, and we conclude, another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned.
-In their essay Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ is put into a  one-to-one correspondence with the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system.
-This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models.
-However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in literature does entail several conundrums that should be avoided---cf.\  \cite{Halvorson:2019} for details.
-To this regard, in fact, HS do not clearly spelled out what they mean when it is asserted that $\psi$ ``represents'' a real property of physical systems.
-Alternatively, to claim that $\psi$-ontic models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered: for instance it is not specified in which way $\psi$ represents the ontic state, is it a direct or indirect representation?
-If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$?
-To these and other issues Harrigan \& Spekkens do not reply.}
-
-\textcolor{magenta}{Therefore, not only their categorization is shown to be empty, but also it does not offer a rigorous philosophical ground where to cast the interpretational debate.}
+As a final note, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models.
+To our knowledge, however, none of them is able to reproduce all results from quantum statistical mechanics (e.g.
+derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g.
+derive the Holevo bound).
+Since \textcolor{magenta}{these} are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.
 

### 2023-10-15 Articles/2021-HarriganSpekkens/Submissions/FoP/response1.tex
Reviewed Andrea's changes with Christine

@@
 \textbf{Authors:} The referee highlights an interesting point in this comment.
-In the mentioned work Hance, Rarity and Ladyman criticize the assumptions on which Harrigan \& Spekkens classification rests on, and show that the dichotomy between $\psi$-ontic and $\psi$-epistemic models is artificial.
+In the mentioned work Hance, Rarity and Ladyman criticize the assumptions on which the Harrigan \& Spekkens classification rests, and show that the dichotomy between $\psi$-ontic and $\psi$-epistemic models is artificial.
 In the revised manuscript we added a new paragraph at the end of section II mentioning this paper and how our argumentative strategy is complementary w.r.t.\ that employed by Hance, Rarity and Ladyman:
@@
 
-\textcolor{magenta}{While we agree with the main messages of these critical responses to the HS framework, the argument offered in this essay follows a different and complementary strategy.
+\textcolor{magenta}{While we agree with the main messages of these critical responses to the HS framework, the argument offered in this paper follows a different and complementary strategy.
 In fact, instead of criticizing the assumptions used by Harrigan \& Spekkens, we consider them valid (for the sake of the argument) and derive a no-go theorem for $\psi$-ontic models directly from them, as we are going to show in the next section.}
 
-The result obtained by Hance, Rarity and Ladyman is not in contrast with ours since both essays aim at showing the inadequacy of the HS in order to correctly classify interpretations of quantum mechanics.
+The result obtained by Hance, Rarity and Ladyman is not in contrast with ours since both works aim at showing the inadequacy of the HS framework in order to correctly classify interpretations of quantum mechanics.
 However, while it can be true that  the ``condition of non-overlapping can be interpreted as a sufficient, but not necessary for an ontic description'' as the referee correctly says, in our manuscript we consider the HS definitions as true in order to derive a no-go theorem for $\psi$-ontic directly from them.
@@
 \textcolor{blue} {The convex space formed by the mixtures is radically different [34].
-There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics [35-36].
+There is also debate as to whether proper and improper mixtures are distinguishable in the context of quantum mechanics [35-36].
 }
@@
 \textbf{Authors:} The reviewer raises a very important point that was not adequately discussed in the original version of the manuscript.
-In the revised version of the paper we added a new paragraph in the conclusion section saying that the HS framework should not be employed in order to assess the viability of interpretations of the quantum formalism---neither to claim that a certain reading of QM is empirically adequate, not to rule out a specific view:
+In the revised version of the paper we added a new section on interpretations saying that the HS framework should not be employed in order to assess the viability of interpretations of the quantum formalism---neither to claim that a certain reading of QM is empirically adequate, nor to rule out a specific view:
 
-\textcolor{magenta}{Referring to this, two final remarks are in order: on the one hand, our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory.
+\textcolor{magenta}{Our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory.
 Indeed, many scholars use the PBR result to vindicate the empirical adequacy of a given framework---especially those in which the quantum state is somehow supposed to represent real physical objects---or to rule out a specific reading of the quantum formalism.
-Here we warn physicists and philosophers that no argument resting on Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.}
+Here we warn physicists and philosophers that no argument resting on the Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.}
 
-\textcolor{magenta}{To this regard, however, one may argue that our result cannot possibly be correct because there are actual examples of $\psi$-ontic models.
-To our knowledge, however, none of them is fully empirically equivalent to quantum theory, i.e.\ is able to reproduce all results from quantum statistical mechanics (e.g., derive the Boltzmann distribution from entropy maximization) or quantum information theory (e.g., derive the Holevo bound).
-Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.
-In sum, we not only agree with Hance, Rarity and Ladyman in saying that HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:}
+\textcolor{magenta}{Therefore, we not only agree with Hance, Rarity and Ladyman in saying that the HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:}
 \begin{quote}
-\textcolor{magenta}{\textbf{If a certain interpretation of the quantum formalism is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}}
+	\textcolor{magenta}{\textbf{If a certain interpretation of the quantum formalism\footnote{\textcolor{magenta}{As for instance de Broglie-Bohm theory, Everett's relative state formulation, relational quantum mechanics or the many-worlds perspective \emph{etc.}.}} is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}}
 \end{quote}
@@
 
-Another general problem that we see---similar to that mentioned by referee 2---is that the notions of ``ontic'' and ``epistemic'' have always been liberally employed when interpreting quantum theories, but now it is commonly believed that HS provide a set of formal definitions capturing the meaning of such terms.
-In this way, many authors think to substantiate their views with the aid of HS classification.
-We believe that it will be very hard to deconstruct such an habit, however, our essay can be another step to achieve this goal.
-We explicitly discussed a paper of Hance, Rarity and Ladyman in section II, where this issue is openly addressed.
-For more details cf.\ the response to the fourth comment of referee 1.
+\textcolor{magenta}{Another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned.
+In their paper Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ fully identifies the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system.
+This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models.
+However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in the literature entails several conundrums that should be avoided---cf.\  [35] for details.
+With regard to this, the HS framework does not clearly spell out what is meant by the fact that $\psi$ ``represents'' a real property of physical systems.
+In other words, to claim that in $\psi$-complete models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered.
+For instance, it is not specified in which way $\psi$ represents the ontic state: is it a direct or indirect representation?
+If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$?
+On these and other issues the Harrigan \& Spekkens framework does not give guidance.}
 
-\item \textbf{Reviewer 2:} - (5)    Last but no least.
-There is usually a very rough reasoning when the dichotomy psi-ontic vs psi-epistemic is posed.
-The discussion is much more subtle.
-There is a very nice article by Hans Halvorson posing the discussion in a more sophisticated way (To Be a Realist about Quantum Theory).
-\vspace{2mm}
+\textcolor{magenta}{Therefore, not only is their categorization shown to be empty, but also it does not offer a rigorous philosophical ground on which to cast the interpretational debate.}
 
-\textbf{Authors:} Again, we agree with the referee's comment.
-We firmly believe that the philosophical background of HS is problematic, as one of us argued in [10].
-Following the suggestion we added another final paragraph to the conclusion section highlighting that the realist view about the quantum state---its ``onticity''---contained in $\psi$-ontic models is poorly defined:
+\noindent We did not include a more elaborated discussion of the latter point since the main focus of the paper is to prove the technical statement that $\psi$-ontic models cannot fully reproduce quantum theory.
+However, we will address the philosophical implications of our analysis in another philosophically oriented paper that is in preparation.
+We have also summarized the above discussion in the conclusion:
 
-\textcolor{magenta}{On the other hand, and we conclude, another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned.
-In their essay Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ is put into a  one-to-one correspondence with the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system.
-This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models.
-However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in literature does entail several conundrums that should be avoided---cf.\  \cite{Halvorson:2019} for details.
-To this regard, in fact, HS do not clearly spelled out what they mean when it is asserted that $\psi$ ``represents'' a real property of physical systems.
-Alternatively, to claim that $\psi$-ontic models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered: for instance it is not specified in which way $\psi$ represents the ontic state, is it a direct or indirect representation?
-If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$?
-To these and other issues Harrigan \& Spekkens do not reply.}
+\textcolor{magenta}{The direct consequence of this is that the HS categorization should not be employed as a tool to classify quantum interpretations or to draw philosophically sound conclusions on the nature of quantum mechanics.}
 
-\noindent We did not included a more elaborated discussion of the latter point since the main focus of the paper is to prove the technical statement that $\psi$-ontic models cannot fully reproduce quantum theory.
-However, we will address the philosophical implications of our analysis in another philosophically oriented paper that is in preparation.
 \end{enumerate}
@@
 
-\clearpage
-\bibliographystyle{apalike}
-\bibliography{../bibliography}
 \end{document}

### 2023-10-17 Proceedings/2023-QIP/QuantumMeasure.tex
QIP: Finally getting somewhere with QIP proceedings

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
%\documentclass[10pt,twocolumn,letterpaper]{article}

\usepackage{assumptionsofphysics}
\usepackage{tikz}
\usepackage{breakurl}

%\usepackage{amsmath}
%\usepackage{mathrsfs}
%\usepackage{amsfonts}
%
%\usepackage{graphicx}
%\usepackage{hyperref}
%\hypersetup{
%	colorlinks=true,
%	citecolor=blue,
%	urlcolor=blue,
%	linkcolor=blue
%}
%\urlstyle{same}
%\frenchspacing
%
\def\>{\rangle}
\def\<{\langle}
\DeclareMathOperator{\erf}{erf}

\begin{document}

\title{How quantum mechanics requires non-additive measures}
\author{Gabriele Carcassi}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\author{Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
\end{abstract}

\maketitle

\section{Introduction}

Classical probability and classical measure theory does not work.

Alternatives: negative probability, contextual probability, "quantum measure theory", focus on convex space of mixtures.

Interpretation problems (what is a negative probability), and it does not really answer the questions: why does quantization require different contexts? What exactly is quantization? Is there only a way to do it?

Measure theory is also used to quantify the number of states. Statistical mechanics. How are quantum states quantified?

Classical mechanics, agreement between thermodynamic entropy, Gibbs/Shannon entropy and logarithm of the count of states. Two versions. Want something similar in quantum mechanics. Find that is non-additive.

We find that requiring a quantized measure of states leads to non-additivity as well. This opens up a new line of research.

Index:

- Measure theory and quantifying states
(Quantum mechanics does not have a measure of states)

\section{Measure theory and quantifying states}

Much of the focus in quantum foundations lies in the fact that quantum mechanics does not follow the rules of classical Kolmogorov probability. Given that measure theory is typically used to formalize classical probability, numerous efforts start to find an appropriate generalization of measure theory that...The investigation of the use measure theory has been predominantly tied to the Given that this are typically expressed as

However, measure theory is not used only to characterize probability. More in general, measures are used to determine the size of sets. Even in probability theory, a probability measure assigns probability to sets, and to associate probability to points one needs two measures: the probability measure $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$. Given a set $U$, the ratio $\frac{p(U)}{\mu(U)}$ given us the average probability per unit size. As we shrink $U$ to a single point $x$, we will find the probability associated to that point. Since we are interested in physical theories, the measure $\mu(U)$ quantifying the number of states within the region $U$.

In the discrete case, the quantifying measure is simply the counting measure: the cardinality of the set. Therefore for a single point we simply have $\frac{p(\{x\})}{\mu(\{x\})} = \frac{p(\{x\})}{1} = p(x)$, which corresponds to the probability associated to the state $x$. The counting measure is fully determined by two properties: countable additivity, a defining property of measures, and that each states counts as one.

In the continuous case, things are a bit different. If $X$ is a manifold, a set of local coordinates induce a Lebesgue measure, which can be used to quantify the number of possible value combinations. Note that a single state has measure zero, unlike the discrete case. The Lebesgue measure is defined by two properties: countable additivity and that the size of parallelepipeds corresponds to the volume expressed by the product of the difference of coordinates. Naturally, this is a problem in physics given that the quantification of the number of states should be the same for all observers.

What most miss, is that the structure of classical phase space (i.e. symplectic structure) is exactly the structure that allows to define a quantifying measure for states that is invariant under change of observers.(TODO cite book, two phil papers) Given a change of space-time variables $q^\alpha$, conjugate momentum change covariantly, which means volumes of phase space are invariant. The role of quantifying measure in classical mechanics, then, is assigned to the Louisville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e. canonical coordinates). The amount of probability for each state is given by the Radon-Nikodym derivative $\frac{dp}{d\mu}$, which gives us the probability density.

Note that, in both cases, the quantifying measure is potentially unbounded, as the state space can have potentially infinitely many states, as it is the case for one spatial degree of freedom. Therefore the quantifying measures are implicitly different from probability measures, and they require specific attention.

While quantum mechanics does not obey classical probability, we do use of classical probability measure and quantifying measures in some cases. Measurement outcomes are described in terms of classical probability. If we have a discrete observable, like the energy level of a harmonic oscillator, we would use the counting measure; if we have a continuous observable, like position, we would use the Lebesgue measure associated to the variable. Preparations are also in terms of classical probability, and we also use both types of quantifying measures. For example, the maximally mixed state for a qubit can be written as $\frac{1}{2} | 0 \> \< 0 | + \frac{1}{2} | 1 \> \< 1 | $ or as a uniform mixture over all possible states $\int_\mathcal{H} \frac{1}{4\pi} | \psi \> \< \psi | d\Omega$.

Note how, in both cases, the quantifying measure is selected by the type of preparation or measurement, even though the state space is the same. But this means that these quantifying measure cannot be possibly quantifying the number of states as they would give contradicting results. This makes the problem rather crisp: the space of preparations and measurement outcomes are both classical, yet they cannot be connected by classical means. That is, we cannot simply put a classical probability of transition between the two because we are possibly connecting rather different quantifying measures.

To sum up, the typical foundational question is: how do we generalize probability measures for quantum mechanics? We are asking something different: how do we generalize measures that quantify the number of states in quantum mechanics?

\section{Measure theory and quantifying states}

While the standard structure of quantum mechanics does not have an answer to our previous question, statistical mechanics and quantum information theory have an implicit one. Statistical mechanics tells us that the thermodynamic entropy of a system, which is a physically meaningful quantity that has numerous empirical consequences, corresponds to the Shannon entropy for a classical discrete system, the Gibbs entropy\footnote{By Gibbs entropy we mean the differential Shannon entropy $-\int_X \rho \log \rho d\mu$ calculated over phase space using position and momentum (i.e. canonical coordinates).} for a classical continuous system, and the von Neumann entropy for a quantum system.

There is another expression for entropy, often referred to as the fundamental postulate of statistical mechanics, which is the logarithm of the count of microstates: $\log \mu(U)$.\footnote{Note that the Boltzmann constant is left out, as we will be always measured in bits or nats. Physically, the Boltzmann constant is needed to define temperature in Kelvin. tatistical mechanics can be equivalently formulated in terms of beta $\beta = \frac{1}{k_B T}$.TODO: cite the unit definition paper} In classical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Louisville volume in the continuous case. The two expressions for entropy are linked as the Shannon/Gibbs entropy over a uniform distribution matches the logarithm of the measure of the support. That is, if $\rho_U$ is a uniform distribution over $U$, then $S(\rho_U) = \log \mu(U)$.

In quantum mechanics, as we said before, we do not have a measure to quantify states. There is, however, a somewhat related expression: the maximum entropy attainable by a quantum system described by a Hilbert space $\mathcal{H}$ is the dimensionality $\dim_{\mathbb{C}}(\mathcal{H})$ of the space.

A similar connections can be established in information theory. The Shannon/Gibbs/von Neumann entropy quantifies the minimum number of bits required on average to send a message from a given source.\footnote{For a more general characterization of entropy that works in statistical mechanics, information theory and other contexts, and a more precise discussion of the continuous case, see TODO: cite our paper.} This, again, is a quantity that leads to empirical consequences. If we have $n$ bits at our disposal, we have $2^n$ possible combination, possible messages, that can be encoded. Therefore the logarithm of the number of possible messages corresponds to the entropy. Each different message can be encoded into a different state, therefore $\log \mu(U)$ tells us the number of bits that can be encoded using the states in that region. This will exactly correspond to the number of bits required by a source characterized by a uniform probability distribution over that number of symbols.

While the above discussion does not cover all technical details, the only point we want to make is that the link $S(\rho_U) = \log \mu(U)$ is fully justified by both statistical mechanics and information theory. It is not a capricious request, but rather a necessary requirement stemming not from just one, but from two disciplines that have empirical consequences. Therefore if quantum mechanics does not directly provide us with a measure to quantify states, we can use what it does provide, the von Neumann entropy, to define and calculate said measure.

The simplest case is when $U = \{ \psi \}$ is a single state. In this case, we have $S(|\psi\>\<\psi|) = 0$ and therefore $\mu({\psi}) = 2^{0} = 1$. Note that this is the same as the classical discrete case.

The next case is when $U = \{ \psi, \phi \}$ is the set of two states. A uniform distribution over two states will correspond to the mixed state
\begin{equation}
	\rho = \frac{1}{2} |\psi\>\<\psi| + \frac{1}{2} |\phi\>\<\phi|.
\end{equation}
The entropy of said state will depend on the probability $p=|\<\psi | \phi \>|^2$ in the following way
\begin{equation}
	S(\rho) = - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} 
	- \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2}.
\end{equation}
If the states are orthogonal, then $p=0$, and $S(\rho) = 1$ therefore $\mu(U) = 2^1 = 2$. In this case we do have that $\mu(\{ \psi, \phi \}) = \mu(\{\psi\}) + \mu(\{\phi\})$, and therefore $\mu$ is additive. We can also see that if $U$ is an orthonormal basis, the von Neumann entropy $S(U) = \log |U|$ is the logarithm of the cardinality of $U$. Therefore the quantifying measure $\mu(U) = |U|$ becomes the counting measure. But this is a special case that corresponds to the highest entropy reachable with a mixture. In all other cases the entropy will be lower, and therefore, in general, $\mu(\{ \psi, \phi \}) \leq \mu(\{\psi\}) + \mu(\{\phi\})$. In other words, the measure $\mu$ is not additive.\footnote{Technically, measures are defined to be additive, therefore $\mu$ is not a measure, but a set function. However, this term is less widespread, so we will keep using ``measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).}

To sum up, the physically motivated way to quantify the number of states in quantum mechanics is not additive. When counting states in quantum mechanics, $1+1 \leq 2$.

\section{Non-additivity and contextuality}

While unusual claims are common in quantum mechanics, we want to have a clear, physically tenable, reason as to why the quantifying measure for quantum states must be non-additive. In the context of a spin 1/2 system, for example, why is it that $|z^+\>$ and $|z^-\>$ count as two states while $|z^+\>$ and $|x^+\>$ count as less than two?

Since we saw that the quantifying measure is indeed additive for a set of orthogonal states, this should make us think about contextuality: the impossibility to assign outcomes to all potential measurement of a quantum system. We have, in fact, derived that it is impossible to define an additive quantifying measure on all potential states, but we can do so on the outcomes of one measurement. The ability/inability to define additive measures is exactly the difference between classical/non-classical probability. We are definitely onto something.

While contextuality is often presented as a somewhat abstract, almost metaphysical, property of quantum mechanics, we have a rather more down-to-earth view. The selection of the measurement means a difference in physical process of either the preparation or measurement device. Choosing which direction of spin to prepare or measure means, at some point, orienting some polarizer or magnetic field. Something is changed in the physical world, not just on the system, but around the system, on its environment. If Amanda claims that a particle is in the $|z^+\>$ state, then Boris is able to infer something about the preparation/measuring device.\footnote{This is not at all particular to quantum mechanics. We can study the motion of a cannonball here on earth, but not on the surface of the sun, as the cannonball would be immediately vaporized. The mere claim that a system exists implies that the boundary conditions necessary for that system to exist are satisfies. In this sense, all physics is contextual.} Therefore we can understand the context simply as those conditions on the environment that are necessary on the system to be found in that particular state. A set of orthogonal states, then, has the property that they can be found in the same external conditions.

Given that we have, in general, a correlation between system and environment, it is natural to understand why the contextuality leads to a non-additive measure. Two states, in fact, should be counted as two only if they lead to a change in the system. That is, \textbf{states should be quantified all-else-being-equal}. While two states from different contexts are indeed two distinct states, stating we transitioned from one to the other does not give us information only about the system: it tells us something about the environment as well. The degree of incompatibility of observables, then, ultimately measure the degree of change when realizing the context physically.

So we have a clear, physically motivated reason as to why contextuality is linked to a non-additive measure. States should be quantified at-all-else-being-equal, meaning that possible correlations with the environment have to be removed.

\section{Non-additivity and quantization}

While we have clarified what it means for the quantifying measure to be non-additive, and how this is tied to non-contextuality, we haven't shown why it should be linked to quantization. In fact, what exactly is quantization?

Before trying to answer that question, let's make a direct comparison between the quantifying measures for the three cases: the counting measure $\mu_d$ for the classical discrete case, the Louisville measure $\mu_c$ for the classical continuous case and the newly defined quantum measure $\mu_d$ for the quantum case. This requires a space where all three can be defined, which fortunately exists: the surface of a sphere. First, we want to compare the behavior on a single point. The counting measure returns one, and so does the quantum measure. The Louisville measure, however, returns zero. Second, we want to compare the behavior on a finite region, an open set. The Louisville measure returns the solid angle, which is always going to be finite as it must be less than or equal to $4\pi$. The von Neuman entropy of a mixed state over a two state system is at most one, and therefore the measure must be less than or equal to $2$. However, given that an open set on a sphere has infinitely many point, the counting measure over a finite region will be infinite. Lastly, we want to compare the additivity. Both classical measures are additive, while the quantum measure is not.

Let us list the three properties we have mentioned:
\begin{enumerate}
	\item single states count as one
	\item finite continuous regions correspond to finite information (and therefore finite state quantification)
	\item state quantification is additive on disjoint regions.
\end{enumerate}
These seem very reasonable properties to assume for a quantifying measure of states. \textbf{Unfortunately, these three conditions are incompatible}. If each state counts as one, and the measure is additive, then a finite continuous region must be infinite as it contains infinite points. \textbf{Of the three conditions, we can only pick two}. This is, in fact, what each measure does. The counting measure rejects the second property. The Louisville (and Lebesgue) measure discards the first. The quantum measure sacrifices the third.

So, in the light of this, what is quantization? If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see TODO libro} If the regions are large enough, the Louisville (or Lebesgue) measure will work. But as we shrink a region we will encounter, given the additivity, regions that contain less than one state, which does not make sense. Low value of the quantifying measure correspond to low values of entropy and measure one corresponds do zero entropy. Regions with measure less than one correspond to negative thermodynamic entropy, which also does not make sense. In fact, classical statistical mechanics does fail in those regimes: the classical equations of state for an ideal gas to predict negative entropy, which is fixed by the Fermi and Bose statistics.

Quantization, then, is fixing the counting measure so that it is bounded from below by one (and the entropy by zero): each state counts as one. However, this comes at a price: non-additivity. This non-addivity is ultimately why classical probability has to fail as well, given that additivity is one of the axioms. It is also the same reason as why quantization implies contextuality: the non-addivity means that some states cannot be prepared at-all-else-being equal.
%(https://en.wikipedia.org/wiki/Sackur%E2%80%93Tetrode_equation)

In our view, this provides good conceptual reasons to understand the failure of classical mechanics, the need for the departure from classical ideas and the need for quantum theory. We believe this is already sufficient reason to start a research program to explore these ideas fully. However, since most physicists are not interested in this type of ``clean-up work'', let's see this is, in our mind, critical for the development of future physical theories.

\section{Implication for space-time quantization}

As we showed before, this line of reasoning gives us a good conceptual understanding of what quantization is: the quantifying measure for states must have a unitary lower bound. If we were to quantize space-time, what is it that we are quantizing?

The most fundamental theories that are available to us are field theories. In a field theory, the state is described by the value of the fields at each point. At each point, the value of the field is independent from the others: the degrees of freedom form a continuum.\footnote{If we restrict ourselves to continuous fields, then we only need to specify the value of the fields on a countable dense subset (i.e. rational values of the coordinates). Though these kind of details do not change the discussion: a Lebesgue measure on the rationals have the same characteristics of the one extended to the reals.} We can therefore pose the question: how do we quantify the degrees of freedom?

We have again the same problem as before: measures on the continuum are not invariant under generic coordinate transformations. We do have a sense, however, that if we double the spatial volume, we double the number of degrees of freedom, and spatial volume is an invariant. Therefore we can use a measure of volume constructed from the metric tensor. This would be the equivalent of the Louisville measure we got from phase space, and we end up with an equivalent problem.

Since the measure is finite for finite regions and is additive, as we shrink the region, we will find regions with less than one degree of freedom. This does not make sense. If we want to quantize space, then, we need to put a unitary lower bound to the measure. But now our measure cannot be additive: what does it mean that two degrees of freedom count less than two?

Before answering that question, let's add time. We said that if we double the spatial volume, we double the number of degrees of freedom. Suppose we have a space-time region and we double the time, are we doubling the number of degrees of freedom? Well, we certainly doubling the support of the function we have to specify for the field configuration, but we wouldn't consider that doubling the degrees of freedom. Intuitively, we just think of those as being the same degrees of freedom that are getting different values. This is because we expect the future and past values to be linked by some dynamical equations, so the future values would be linked, in some way, to the past values. That is, the past and future value of the same field at some point are not independent.

So, what does it mean that two degrees of freedom count less than two? It could simply mean that they are not independent. That is, the values of the field at two distant points can be chosen arbitrarily, but as the points get sufficiently close, this would not be the case. This would make sense.

Note that the lack of independence of the field values in close-by regions may mean that high frequencies of the fields are damped: the values have to become closer. The quantized quantifying measure of the degrees of freedom would effectively provide a cutoff. This may address some of the open problem in quantum field theories.

In our view, the arguments laid here highlight an undeniable problem and seem to offer a natural step to take in line with the change from classical to quantum theory. While the development of a quantum theory of gravity is not our primary interest, the mathematical tools needed to address the above problem are of the same nature as the one we propose to develop a non-additive measure theory to quantify quantum states.

\section{Open questions and future work}

Is this the only quantized measure? What are the necessary conditions? Can we define a corresponding non-additive probability measure in the same way, so that a non-additive density can be defined? What type of theory of integration should lead to? How can this be extended to the number of d.o.f.?

- Conclusion

\section{Overview}

We will first present an overview, which serves both as a summary and as a guide for rest of the paper. Each subsection will be expanded into a full section with the full technical details.

\subsection{Measure theory in physics}

The first task is to understand the link between physics and measure theory. We start with a set $X$, which formally represents a set of physically distinguishable cases. In our case, $X$ will either be a set of states (e.g. a symplectic manifold for classical mechanics; the set rays of a Hilbert space of quantum mechanics), or the set of possible values for either a preparation procedure or a measurement outcome (e.g. the numeric value of position and momentum for a classical system; the value of the azimuthal and polar angles for spin $1/2$ system; $\pm 1/2 \hbar$ for a spin $1/2$ measurement along a particular direction).

A measure is a map $\mu : \Sigma_X \to [0,+\infty]$ that given a subset $U \subseteq X$ returns a non-negative, possibly infinite, value $\mu(U)$. The family $\Sigma_X$ of subsets of $X$ on which the measure is defined is closed under complement, countable union and intersection, which is the correct closure for experimental procedures. A defining feature of measures is that they are additive, meaning that
\begin{equation}\label{additive_measure}
	\begin{aligned}
		&\mu(U_1 \cup U_2) = \mu(U_1) + \mu(U_2) \\
		&\forall U_1, U_2 \in \Sigma_X \text{ such that } U_1 \cap U_2 = 0.
	\end{aligned}
\end{equation}

In our case, a measure will keep track of one of two things: either it quantifies the number of physically distinguishable cases are there in the subset $U$, in which case we will denote it as $\mu(U)$ and call it \textbf{quantifying measure}, or it quantifies the probability that the statement $x \in U$ is true, in which case we will denote it as $p(U)$ and call it \textbf{probability measure}. For example, if $U$ is a subset of phase space $S$, $\mu(U)$ will quantify the number of states in the set $U$. If $U$ is a subset of the possible preparation for position and velocity $\mathbb{R}^2$, $p(U)$ will correspond to the probability that the initial system configuration is in $V$.

%There are two crucial aspects in the use of measures, one typically ignored by mathematicians and the other typically not understood by physicists. The first is that of units: while all probability measures always have units of probability, those of quantifying measures depends on the space itself. For example, the space of possible values of position and velocity along a particular direction is mathematically the same, the real line $\mathbb{R}$, and the measure $\mu$ in both cases will correspond to the Lebesgue measure. However, the values are incommensurable , but the units will be different. Neglecting the difference  essentially none. This a major contribution to the second problem, which is that too many physicists consider the continuum the same as the discrete, but ``with more points''. The two cases, in fact, are qualitatively different, but the focus of measure theory on sets, instead of points, allows to treat these two different cases within the same framework. We will see that understanding the difference and similarities between them is crucial as quantum mechanics is essentially a third case that bridges between the two.

Given two measures, $\mu$ and $\nu$, we can define the ratio between the two $\frac{\nu(U)}{\mu(U)}$. If the space is not discrete (e.g. a manifold), we can also define the derivative $\frac{d\nu}{d\mu}$. For these to be well defined, $\nu(U)$ must be zero whenever $\mu(U)$ is zero. Therefore, taking the derivative $\frac{dp}{d\mu}$ between a probability measure and a quantifying measure is always well defined, as it would not make sense to assign a non-zero probability on a set that has zero possible cases. If we have a map that links two spaces, we can take the derivative between two quantifying measures. For example, given a map $(x, v) \to X$ that given a value of position and velocity identifies a classical state within the suitable symplectic manifold, we can define $\frac{d\mu}{d\nu}$ that tells us how many states are there per unit of position time velocity.

Given a probability measure $p$ and a quantifying measure $\mu$, we can define the entropy $S_{\mu}(p)$. If the space is discrete and $\mu$ is the counting measure (i.e. it returns the cardinality of the set $\mu(U) = |U|$), then we simply have the Shannon entropy
\begin{equation}\label{discrete_entropy}
	S_{\mu}(p) = - \sum p_i \log p_i.
\end{equation}
If the space is $\mathbb{R}^n$ and $\mu$ is the Lebesgue measure, we have the Shannon differential entropy
\begin{equation}
	S_{\mu}(p) = - \int \rho \log \rho d\mu
\end{equation}
where $\rho = \frac{dp}{d\mu}$. If the probability distribution is uniform over a region $U \subset X$, then the entropy is equal to
\begin{equation}
	S_{\mu}(p) = \log \mu(U)
\end{equation}
in both the discrete and continuous case.

There are a couple of crucial things. First, the mathematics does not keep track of physical units, missing critical physical elements. Second, this difference changes the discrete from the continuous case rather radically. Third, these difference become rather important when trying to analyze how measure theory is used in quantum mechanics. Let us see how it works in each case.

\subsection{Discrete classical case}

In the discrete classical case, the state space $X$ has at most countably many elements, it has a discrete topology and $\Sigma_X$ contains all possible subsets. The quantifying measure associated to the space $X$ is the counting measure and any probability measure can be written as
\begin{equation}
	p(U) = \sum p(\{x_i\}) = \sum p_i,
\end{equation}
the sum of the probability assigned to each element. The entropy is given by eq. \ref{discrete_entropy}.

If have a set of labels $L$ to identify the states, it will also be (topologically) discrete and associated to the counting measure. If $x(l)$ is the map between label and states, then $\mu_L(U) = \mu_X(x(U))$. In other words, quantifying the states or quantifying the labels is exactly the same thing. By the same token, defining the probability measure on the states or on the label will be the same: $p_L(U) = \sum p_L({l_i}) = \sum p_X(x(l_i))$.

In the discrete case, then, counting possible preparations, states or measurement outcomes is the same. Assigning probabilities on preparations, states or measurement outcomes is the same. Computing entropy on preparations, states or measurement outcomes is the same.

\subsection{Continuous classical case}

In the continuous classical case, the state space $X$ is classical phase space (i.e. a symplectic manifold), and $\Sigma_X$ is formed by the Borel sets. States are counted through the Liouville measure
\begin{equation}
	\mu(U) = \int_U \omega^n 
\end{equation}
where $\omega$ is the symplectic form and $n$ is the number of degrees of freedom. Note that volume will become zero for a single point, therefore, for continuity, there will be regions that have ``less than one state''. Note the Liouville measure is invariant under coordinate transformations: the symplectic structure is exactly the structure required to provide a count of states $\mu(U)$ that is observer independent. This is not a coincidence.

If we have a set of variables $\xi^a$, the measure $\mu_{\xi^a}(U)$ will be the volume given in the product of all units of $\xi^a$. This means that, in general, $\mu_{\xi^a}(U)$ depends on the choice of variables and $\mu_{\xi^a}(U) \neq \mu(x(U))$. The two measures will coincide if the chosen variables are position and conjugate momentum (i.e. canonical coordinates), but will not, for example, if position and velocity are used. This is a key difference from the discrete case: the quantifying measure on preparations, states and measurement outcomes is not the same. Moreover, only the quantification of states is required to be observer invariant.

If we have a map between preparations, states and outcomes that is bijective, the probability measure will transfer like in the discrete case. Therefore $p_{\xi^a}(U) = p_X(x(U))$ for any region $U$ charted by any set of variables $\xi^a$. However, the probability densities $\frac{dp}{d\mu}$ in each space will be different because the respective quantifying measures are different. Intuitively, the density will be expressed in units of probability over units of preparations/states/outcomes. Again, the density over states is the only one required to be observer invariant.

Given that the differential entropy is calculated in terms of the probability density, this will also in general depend on the quantifying measures and their units. The differential entropy on states will be the only one that is guaranteed to be observer invariant, and it is the one that will correspond to the physical entropy. The entropies on the preparations and measurement outcomes, instead, must be understood as characterizing information about specific variables.

\subsection{Quantum case}

The situation in quantum mechanics is more complicated. The state space $X$ is the projective space of a Hilbert space $P[\mathcal{H}]$. In the simplest case, a two dimensional system, it will correspond to the Bloch sphere. However, the space of the possible outcomes of a measurement is rather more restricted: an orthonormal bases for a projection-valued measurement. In the case of a two dimensional system, that means two discrete states.

The space of measurements, then, is not the same as the space of states. If the measured quantity is discrete (e.g. spin along a direction), the quantifying measure would be the counting measure. If the measured quantity is continuous (e.g. position), the quantifying measure is the Lebesgue measure. The way that we quantify states, then, is necessarily different from the way we quantify measurement outcomes.

The space state $X$ can be shown to be a symplectic manifolds, therefore it allows a Liouville measure, like on the continuous case. For the Bloch sphere, this would simply be the solid angle corresponding to a region. Given that we should have, at least in line of principle, a way to prepare any possible quantum state, we are going to have a bijective map between possible preparation procedures and quantum states.

At this point, it would seem we have two components. One is the that the space of both preparation and measurement is $P[\mathcal{H}]$

Something similar happens to the probability measure. Suppose, in fact, that we prepare a two dimensional system with a uniform distribution over the entire Bloch sphere. We would have $p(\psi) = 0$ for any $\psi$ given that we have a uniform distribution over the continuum. However, when

\section{Measure theory in physics}

First, we briefly review the mathematical definition of a measure, so that we can understand how the formal definition maps to physical concepts. Given a set $X$ and a $\sigma$-algebra $\Sigma_X$, a measure is map $\mu : \Sigma_X \to [0,+\infty]$.

Definition of measure theory and use in physics.

Definition of Shannon entropy.

\section{Discrete classical case}

Classical discrete, entropy, count of states.

\section{Continuous classical case}

Classical continuum and Shannon/Gibbs entropy, count of states in statistical mechanics. Discrete vs continuum. Properties of entropy for discrete continuum. Symplectic structure measures that are invariant under coordinate transformation.

\section{Quantum case}

Quantum entropy, Shannon/von Neumann entropy. Hybrid behavior of entropy. Measure of probability and of possible preparations are the same. Lack of count of states.

\section{Quantized measure}

Recover count of states from uniform distribution. Not additive. Couple of examples. Choose 2. Quantization means having a measure on the continuum for which each state counts one.

Need for quantized measure on space time.

\section{Conclusion}

\bibliography{bibliography}

\newcommand{\pj}[1] {\underbar{$#1$}}

\end{document}

### 2023-10-18 Proceedings/2023-QIP/QuantumMeasure.tex
QIP: first draft

@@
 
-Classical probability and classical measure theory does not work.
+The fact that quantum theory is a non-classical probability theory is well established.
+Quantum probability cannot be recovered assuming the standard axioms of Kolmogorov style probability, and their standard implementation in terms of measure theory.
+Various alternatives have been developed with different goals: some intend to give a more general account of probability, where classical and quantum cases are subcases; some want to offer better tools for calculation; some are developed to answer more interpretative or philosophical questions.
+Arguably the most well-known and used alternative approach is the use the quasi-probability distribution over phase space, either in the Wigner or the Glauber–Sudarshan P representation.
 
-Alternatives: negative probability, contextual probability, "quantum measure theory", focus on convex space of mixtures.
+While the unsuitability of standard measure theory to represent quantum probability is a crucial aspect, this focus seems to have obscured the fact that measure theory plays another crucial in physical theories: it is used to ``count', or better to ``quantify'', states.
+As part of our overall project Assumptions of Physics, we have developed an approach, called reverse physics, which aims to start from the laws and find a minimal set of physical assumptions needed to rederive the laws.
+What we found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.
+The former (i.e.
+the symplectic structure) is the only structure that allows to quantify states, and configurations over independent degrees of freedom, in a way that is observer independent.
+If one imposes deterministic and reversible evolution, that structure has to be preserved in time, leading to Hamiltonian evolution (i.e.
+a symplectomorphism).
+All elements of classical mechanics, including canonical variables, the Lagrangian and the action principle, can be rederived and understood in terms of quantifying the flow of states.
+Given the formal similarity between classical and quantum mechanics, we believe a similar approach is possible for the latter.
+Which raises the question: what is the quantum analogue of the Louisville measure, the measure that quantifies states in the classical theory?
 
-Interpretation problems (what is a negative probability), and it does not really answer the questions: why does quantization require different contexts?
-What exactly is quantization?
-Is there only a way to do it?
+Statistical mechanics and information theory mandate a link between entropy and state quantification, as the Shannon/Gibbs entropy over uniform distributions corresponds to the logarithm of the count of states.
+In the quantum case, we can start with the von Neumann entropy and define a measure that has the same connection.
+What we find is that this measure is non-additive and that its features can help us better understand what quantization is, why it requires contextuality and the failure of classical probability, and why the latter can be recovered separately within each context.
 
-Measure theory is also used to quantify the number of states. Statistical mechanics. How are quantum states quantified?
-
-Classical mechanics, agreement between thermodynamic entropy, Gibbs/Shannon entropy and logarithm of the count of states.
-Two versions.
-Want something similar in quantum mechanics.
-Find that is non-additive.
-
-We find that requiring a quantized measure of states leads to non-additivity as well. This opens up a new line of research.
-
-Index:
-
-- Measure theory and quantifying states
-(Quantum mechanics does not have a measure of states)
+The goal of this article is to propose and articulate the goals of this new line of research.
+It provides the vision for future works on both the mathematical details and the conceptual structure.
+In section II we will show how measure theory is used to quantify elements in a set and how this is needed to assign probability and probability densities to each element.
+In section II we will see how statistical mechanics and information theory mandate a link between entropy and state quantification, which we will use to define the quantum analogue of the Louisville measure.
+We will find that this measure is not additive.\footnote{Technically, measures are defined to be additive, therefore what we find is not a measure, but a set function.
+However, this term is less widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).} In section IV we will see that this non-additivity is linked to non-contextuality.
+In section V we will see that quantization means putting a lower unitary bound on the quantification of states, while keeping a finite value over finite regions.
+We will see that this very directly requires non-additivity and therefore non-contextuality.
+In section VI we will argue that quantization of space-time means using a quantized measure on the quantification of degrees of freedom.
+Lastly, in section VII we will outline a series of goals for a research program that aims to fully characterized the quantized measure.
 
@@
 In all other cases the entropy will be lower, and therefore, in general, $\mu(\{ \psi, \phi \}) \leq \mu(\{\psi\}) + \mu(\{\phi\})$.
-In other words, the measure $\mu$ is not additive.\footnote{Technically, measures are defined to be additive, therefore $\mu$ is not a measure, but a set function.
-However, this term is less widespread, so we will keep using ``measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).}
+In other words, the measure $\mu$ is not additive.
 
@@
 This would make sense.
-
 Note that the lack of independence of the field values in close-by regions may mean that high frequencies of the fields are damped: the values have to become closer.
 The quantized quantifying measure of the degrees of freedom would effectively provide a cutoff.
-This may address some of the open problem in quantum field theories.
+This may provide a natural high energy cutoff for quantum field theories.
 
@@
 
-Is this the only quantized measure?
-What are the necessary conditions?
-Can we define a corresponding non-additive probability measure in the same way, so that a non-additive density can be defined?
-What type of theory of integration should lead to?
-How can this be extended to the number of d.o.f.?
+Having outlined the scope and motivation for this new approach, we now outline the main questions we think need to be addressed.
 
-- Conclusion
+Both the counting measure and the Lebesgue/Louisville measures are unique in some sense.
+For the counting measure, once we define the value for a single point, typically one, the whole measure is determined.
+Similarly, once we have defined a real variable, the measure is determined.
+The Louisville requires a bit more structure, but once the canonical coordinates are chosen, the measure is determined as well.
+Therefore the first question is whether these quantized measure are also unique in some sense.
 
-\section{Overview}
+While we have found some literature on non-additive measures and set functions (CITE), together with modified theories of integration, these seem to concentrate on monotonic measure, that is measure that increase as the set increase.
+Unfortunately, the quantized measure does not have this feature.
+Consider the set $\{ z^+, z^- \}$.
+An equal mixture will return the maximally mixed state which, as we said, would correspond to a measure equal to 2.
+Now consider the set $\{ z^+, z^-, x^+ \}$.
+This will not correspond to the maximally mixed state, and therefore the measure will be less than two.
+The set is bigger, but the measure is smaller.
+Therefore we may need to develop an appropriate theory.
 
-We will first present an overview, which serves both as a summary and as a guide for rest of the paper. Each subsection will be expanded into a full section with the full technical details.
+To achieve the previous point, we will need to understand the necessary conditions for the quantifying measure.
+Some may be derived from the strict concavity of entropy, given the connection between the two.
+Furthermore, note that the relationship between the measure over two points, the entropy of the equal mixture of two states and the probability of transition between two states are invertible.
+Therefore the relationship between two states defines the geometry of the whole state space.
+Therefore there must be further consistency requirements that can be identified.
 
-\subsection{Measure theory in physics}
+Another thing that need to be understand is how the quantized measure behaves with composite systems.
+In the classical case, the measure is defined on the product of the $\sigma$-algebra.
+Ideally, we would like to show that this simple procedure cannot support a quantized measure, and the $\sigma$-algebra of tensor product must be taken.\footnote{Technically it would be the $\sigma$-algebra projective space coming from the tensor product of the Hilbert spaces associated with the individual systems.}
 
-The first task is to understand the link between physics and measure theory.
-We start with a set $X$, which formally represents a set of physically distinguishable cases.
-In our case, $X$ will either be a set of states (e.g.
-a symplectic manifold for classical mechanics; the set rays of a Hilbert space of quantum mechanics), or the set of possible values for either a preparation procedure or a measurement outcome (e.g.
-the numeric value of position and momentum for a classical system; the value of the azimuthal and polar angles for spin $1/2$ system; $\pm 1/2 \hbar$ for a spin $1/2$ measurement along a particular direction).
+The first spaces we should study are the symplectic ones as these are the ones that allow for observer independent densities.
+If we want to extend to a quantization of space-time, however, Riemannian spaces need to be studies as well.
+However, we have found a connection between the two: if the symplectic form is expressed in position and velocity, instead of canonical coordinate, the metric tensor appears as the position-velocity component.
+It then can be understood as defining volumes in position-velocity instead in position-momentum.
+This may allow us to transfer the quantized measure on the tangent bundle and then, hopefully, to the space itself.
 
-A measure is a map $\mu : \Sigma_X \to [0,+\infty]$ that given a subset $U \subseteq X$ returns a non-negative, possibly infinite, value $\mu(U)$.
-The family $\Sigma_X$ of subsets of $X$ on which the measure is defined is closed under complement, countable union and intersection, which is the correct closure for experimental procedures.
-A defining feature of measures is that they are additive, meaning that
-\begin{equation}\label{additive_measure}
-	\begin{aligned}
-		&\mu(U_1 \cup U_2) = \mu(U_1) + \mu(U_2) \\
-		&\forall U_1, U_2 \in \Sigma_X \text{ such that } U_1 \cap U_2 = 0.
-	\end{aligned}
-\end{equation}
+There is therefore much work to be done, but one thing we want to stress: a clear and precise physical idea is driving the development of possibly new math.
+We are not looking for mathematical extensions and trying them to see whether something useful can be achieved.
+We have a goal, quantification of states .
 
-In our case, a measure will keep track of one of two things: either it quantifies the number of physically distinguishable cases are there in the subset $U$, in which case we will denote it as $\mu(U)$ and call it \textbf{quantifying measure}, or it quantifies the probability that the statement $x \in U$ is true, in which case we will denote it as $p(U)$ and call it \textbf{probability measure}.
-For example, if $U$ is a subset of phase space $S$, $\mu(U)$ will quantify the number of states in the set $U$.
-If $U$ is a subset of the possible preparation for position and velocity $\mathbb{R}^2$, $p(U)$ will correspond to the probability that the initial system configuration is in $V$.
-
-%There are two crucial aspects in the use of measures, one typically ignored by mathematicians and the other typically not understood by physicists.
-The first is that of units: while all probability measures always have units of probability, those of quantifying measures depends on the space itself.
-For example, the space of possible values of position and velocity along a particular direction is mathematically the same, the real line $\mathbb{R}$, and the measure $\mu$ in both cases will correspond to the Lebesgue measure.
-However, the values are incommensurable , but the units will be different.
-Neglecting the difference  essentially none.
-This a major contribution to the second problem, which is that too many physicists consider the continuum the same as the discrete, but ``with more points''.
-The two cases, in fact, are qualitatively different, but the focus of measure theory on sets, instead of points, allows to treat these two different cases within the same framework.
-We will see that understanding the difference and similarities between them is crucial as quantum mechanics is essentially a third case that bridges between the two.
-
-Given two measures, $\mu$ and $\nu$, we can define the ratio between the two $\frac{\nu(U)}{\mu(U)}$.
-If the space is not discrete (e.g.
-a manifold), we can also define the derivative $\frac{d\nu}{d\mu}$.
-For these to be well defined, $\nu(U)$ must be zero whenever $\mu(U)$ is zero.
-Therefore, taking the derivative $\frac{dp}{d\mu}$ between a probability measure and a quantifying measure is always well defined, as it would not make sense to assign a non-zero probability on a set that has zero possible cases.
-If we have a map that links two spaces, we can take the derivative between two quantifying measures.
-For example, given a map $(x, v) \to X$ that given a value of position and velocity identifies a classical state within the suitable symplectic manifold, we can define $\frac{d\mu}{d\nu}$ that tells us how many states are there per unit of position time velocity.
-
-Given a probability measure $p$ and a quantifying measure $\mu$, we can define the entropy $S_{\mu}(p)$.
-If the space is discrete and $\mu$ is the counting measure (i.e.
-it returns the cardinality of the set $\mu(U) = |U|$), then we simply have the Shannon entropy
-\begin{equation}\label{discrete_entropy}
-	S_{\mu}(p) = - \sum p_i \log p_i.
-\end{equation}
-If the space is $\mathbb{R}^n$ and $\mu$ is the Lebesgue measure, we have the Shannon differential entropy
-\begin{equation}
-	S_{\mu}(p) = - \int \rho \log \rho d\mu
-\end{equation}
-where $\rho = \frac{dp}{d\mu}$. If the probability distribution is uniform over a region $U \subset X$, then the entropy is equal to
-\begin{equation}
-	S_{\mu}(p) = \log \mu(U)
-\end{equation}
-in both the discrete and continuous case.
-
-There are a couple of crucial things.
-First, the mathematics does not keep track of physical units, missing critical physical elements.
-Second, this difference changes the discrete from the continuous case rather radically.
-Third, these difference become rather important when trying to analyze how measure theory is used in quantum mechanics.
-Let us see how it works in each case.
-
-\subsection{Discrete classical case}
-
-In the discrete classical case, the state space $X$ has at most countably many elements, it has a discrete topology and $\Sigma_X$ contains all possible subsets.
-The quantifying measure associated to the space $X$ is the counting measure and any probability measure can be written as
-\begin{equation}
-	p(U) = \sum p(\{x_i\}) = \sum p_i,
-\end{equation}
-the sum of the probability assigned to each element. The entropy is given by eq. \ref{discrete_entropy}.
-
-If have a set of labels $L$ to identify the states, it will also be (topologically) discrete and associated to the counting measure.
-If $x(l)$ is the map between label and states, then $\mu_L(U) = \mu_X(x(U))$.
-In other words, quantifying the states or quantifying the labels is exactly the same thing.
-By the same token, defining the probability measure on the states or on the label will be the same: $p_L(U) = \sum p_L({l_i}) = \sum p_X(x(l_i))$.
-
-In the discrete case, then, counting possible preparations, states or measurement outcomes is the same.
-Assigning probabilities on preparations, states or measurement outcomes is the same.
-Computing entropy on preparations, states or measurement outcomes is the same.
-
-\subsection{Continuous classical case}
-
-In the continuous classical case, the state space $X$ is classical phase space (i.e.
-a symplectic manifold), and $\Sigma_X$ is formed by the Borel sets.
-States are counted through the Liouville measure
-\begin{equation}
-	\mu(U) = \int_U \omega^n 
-\end{equation}
-where $\omega$ is the symplectic form and $n$ is the number of degrees of freedom.
-Note that volume will become zero for a single point, therefore, for continuity, there will be regions that have ``less than one state''.
-Note the Liouville measure is invariant under coordinate transformations: the symplectic structure is exactly the structure required to provide a count of states $\mu(U)$ that is observer independent.
-This is not a coincidence.
-
-If we have a set of variables $\xi^a$, the measure $\mu_{\xi^a}(U)$ will be the volume given in the product of all units of $\xi^a$.
-This means that, in general, $\mu_{\xi^a}(U)$ depends on the choice of variables and $\mu_{\xi^a}(U) \neq \mu(x(U))$.
-The two measures will coincide if the chosen variables are position and conjugate momentum (i.e.
-canonical coordinates), but will not, for example, if position and velocity are used.
-This is a key difference from the discrete case: the quantifying measure on preparations, states and measurement outcomes is not the same.
-Moreover, only the quantification of states is required to be observer invariant.
-
-If we have a map between preparations, states and outcomes that is bijective, the probability measure will transfer like in the discrete case.
-Therefore $p_{\xi^a}(U) = p_X(x(U))$ for any region $U$ charted by any set of variables $\xi^a$.
-However, the probability densities $\frac{dp}{d\mu}$ in each space will be different because the respective quantifying measures are different.
-Intuitively, the density will be expressed in units of probability over units of preparations/states/outcomes.
-Again, the density over states is the only one required to be observer invariant.
-
-Given that the differential entropy is calculated in terms of the probability density, this will also in general depend on the quantifying measures and their units.
-The differential entropy on states will be the only one that is guaranteed to be observer invariant, and it is the one that will correspond to the physical entropy.
-The entropies on the preparations and measurement outcomes, instead, must be understood as characterizing information about specific variables.
-
-\subsection{Quantum case}
-
-The situation in quantum mechanics is more complicated.
-The state space $X$ is the projective space of a Hilbert space $P[\mathcal{H}]$.
-In the simplest case, a two dimensional system, it will correspond to the Bloch sphere.
-However, the space of the possible outcomes of a measurement is rather more restricted: an orthonormal bases for a projection-valued measurement.
-In the case of a two dimensional system, that means two discrete states.
-
-The space of measurements, then, is not the same as the space of states.
-If the measured quantity is discrete (e.g.
-spin along a direction), the quantifying measure would be the counting measure.
-If the measured quantity is continuous (e.g.
-position), the quantifying measure is the Lebesgue measure.
-The way that we quantify states, then, is necessarily different from the way we quantify measurement outcomes.
-
-The space state $X$ can be shown to be a symplectic manifolds, therefore it allows a Liouville measure, like on the continuous case.
-For the Bloch sphere, this would simply be the solid angle corresponding to a region.
-Given that we should have, at least in line of principle, a way to prepare any possible quantum state, we are going to have a bijective map between possible preparation procedures and quantum states.
-
-At this point, it would seem we have two components. One is the that the space of both preparation and measurement is $P[\mathcal{H}]$
-
-Something similar happens to the probability measure.
-Suppose, in fact, that we prepare a two dimensional system with a uniform distribution over the entire Bloch sphere.
-We would have $p(\psi) = 0$ for any $\psi$ given that we have a uniform distribution over the continuum.
-However, when
-
-\section{Measure theory in physics}
-
-First, we briefly review the mathematical definition of a measure, so that we can understand how the formal definition maps to physical concepts.
-Given a set $X$ and a $\sigma$-algebra $\Sigma_X$, a measure is map $\mu : \Sigma_X \to [0,+\infty]$.
-
-Definition of measure theory and use in physics.
-
-Definition of Shannon entropy.
-
-\section{Discrete classical case}
-
-Classical discrete, entropy, count of states.
-
-\section{Continuous classical case}
-
-Classical continuum and Shannon/Gibbs entropy, count of states in statistical mechanics.
-Discrete vs continuum.
-Properties of entropy for discrete continuum.
-Symplectic structure measures that are invariant under coordinate transformation.
-
-\section{Quantum case}
-
-Quantum entropy, Shannon/von Neumann entropy. Hybrid behavior of entropy. Measure of probability and of possible preparations are the same. Lack of count of states.
-
-\section{Quantized measure}
-
-Recover count of states from uniform distribution. Not additive. Couple of examples. Choose 2. Quantization means having a measure on the continuum for which each state counts one.
-
-Need for quantized measure on space time.
+ are looking to quantify states  We are not introducing something like negative probability, we have a clear understanding of what the physical significance of the generalized measure is.
+The goal is the quantification of states in the contextual case.
 
 \section{Conclusion}
+
+We have seen that, in physics, measure theory is not only used to characterize classical probability but also to quantify states in state spaces.
+Ideally, such measure would have the following three conditions:
+\begin{enumerate}
+	\item single states count as one
+	\item finite continuous regions correspond to finite information (and therefore finite state quantification)
+	\item state quantification is additive on disjoint regions.
+\end{enumerate}
+However, these conditions are not compatible, and only two of the three can be chosen.
+The classical discrete cases discards the second; the classical continuous case discards the first; the quantum case discards the third.
+Quantization, then, can be understood as putting a unitary lower bound to the measure that quantifies states.
+
+The lack of additivity on the whole space makes it impossible to use classical probability on the state space as a whole.
+Physically, the issue is that states must be counted at-all-else-being-equal, feature broken by contextuality.
+But if the states do belong to the same context (e.g.
+they are the output of the same measurement), additivity is recovered.
+This tells us very directly why quantization requires contextuality, and why classical probability works within the same context.
+The fact that single states and and finite continuous regions have both finite measure also tells us why quantum mechanics mixes features of the classical and discrete cases.
+
+We believe that further study and characterization of the non-additive quantized measure will not only bring more clarity in the interplay between the above concept, but it may be necessary for the development of future physical theory.
+Given that, in a field theory, the number of independent degrees of freedom is taken to be proportional to the spatial volume, quantization of space time can be understood as the quantization of the measure that quantifies the indipendent degrees of freedom.
+The development of physically motivated mathematical tools may help in that regard.
 

### 2023-10-22 Proceedings/2023-QIP/QuantumMeasure.tex
QIP: reviewed

@@
 
-The fact that quantum theory is a non-classical probability theory is well established.
-Quantum probability cannot be recovered assuming the standard axioms of Kolmogorov style probability, and their standard implementation in terms of measure theory.
-Various alternatives have been developed with different goals: some intend to give a more general account of probability, where classical and quantum cases are subcases; some want to offer better tools for calculation; some are developed to answer more interpretative or philosophical questions.
-Arguably the most well-known and used alternative approach is the use the quasi-probability distribution over phase space, either in the Wigner or the Glauber–Sudarshan P representation.
+It is well-established that standard measure theoretic tools cannot be used to characterize quantum probability as quantum theory is a non-classical (i.e.
+non-Kolmogorov) probability theory.
+To obviate this problem, various alternatives and extensions to measure theory and probability theory have been developed with different goals.
+Our interest in measure theory, however, is not in its use to capture classical probability.
+Measures play another crucial use in physical theories: ``counting'', or better ``quantifying'', states.
+It is this use that will be the focus of this work.
 
-While the unsuitability of standard measure theory to represent quantum probability is a crucial aspect, this focus seems to have obscured the fact that measure theory plays another crucial in physical theories: it is used to ``count', or better to ``quantify'', states.
-As part of our overall project Assumptions of Physics, we have developed an approach, called reverse physics, which aims to start from the laws and find a minimal set of physical assumptions needed to rederive the laws.
-What we found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.
+As part of our overall project Assumptions of Physics, we have developed an approach, called reverse physics, which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive the laws.
+We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.
 The former (i.e.
-the symplectic structure) is the only structure that allows to quantify states, and configurations over independent degrees of freedom, in a way that is observer independent.
-If one imposes deterministic and reversible evolution, that structure has to be preserved in time, leading to Hamiltonian evolution (i.e.
+the symplectic structure) is the only structure that allows to quantify states and configurations over independent degrees of freedom in a way that is observer independent.
+If one imposes deterministic and reversible evolution, state quantification must be preserved, leading to Hamiltonian evolution (i.e.
 a symplectomorphism).
@@
 In the quantum case, we can start with the von Neumann entropy and define a measure that has the same connection.
-What we find is that this measure is non-additive and that its features can help us better understand what quantization is, why it requires contextuality and the failure of classical probability, and why the latter can be recovered separately within each context.
+What we find is that this measure is non-additive and that its features can help us better understand what quantization is, why it requires both contextuality and the failure of classical probability, and why the latter can be recovered separately within each context.
+Therefore we believe studying this object is useful not only for the above stated goals of reverse physics, but more in general to get a better understanding of the foundations of quantum mechanics.
 
 The goal of this article is to propose and articulate the goals of this new line of research.
-It provides the vision for future works on both the mathematical details and the conceptual structure.
+It provides the overall picture in which future works will add further mathematical detail and physical insights.
 In section II we will show how measure theory is used to quantify elements in a set and how this is needed to assign probability and probability densities to each element.
-In section II we will see how statistical mechanics and information theory mandate a link between entropy and state quantification, which we will use to define the quantum analogue of the Louisville measure.
+In section III we will see how statistical mechanics and information theory mandate a link between entropy and state quantification, which we will use to define the quantum analogue of the Louisville measure.
 We will find that this measure is not additive.\footnote{Technically, measures are defined to be additive, therefore what we find is not a measure, but a set function.
-However, this term is less widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).} In section IV we will see that this non-additivity is linked to non-contextuality.
+However, this term is not widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).} In section IV we will see that this non-additivity is linked to non-contextuality.
 In section V we will see that quantization means putting a lower unitary bound on the quantification of states, while keeping a finite value over finite regions.
@@
 
-Much of the focus in quantum foundations lies in the fact that quantum mechanics does not follow the rules of classical Kolmogorov probability.
-Given that measure theory is typically used to formalize classical probability, numerous efforts start to find an appropriate generalization of measure theory that...The investigation of the use measure theory has been predominantly tied to the Given that this are typically expressed as
-
-However, measure theory is not used only to characterize probability.
-More in general, measures are used to determine the size of sets.
-Even in probability theory, a probability measure assigns probability to sets, and to associate probability to points one needs two measures: the probability measure $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$.
+As we said in the introduction, measure theory is not only used to characterize Kolmogorovian probability.
+More in general, measures are used to define the size of sets.
+Even in probability theory, since probability measures assigns probability to sets of point, to associate probability to points one needs two measures: the \textbf{probability measure} $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$.\footnote{Measure assigns a value to elements to a $\sigma$-algebra that can be thought as a set of experimental procedures.} Since we are interested in physical theories, the measure $\mu$ quantifies the number of states.
 Given a set $U$, the ratio $\frac{p(U)}{\mu(U)}$ given us the average probability per unit size.
-As we shrink $U$ to a single point $x$, we will find the probability associated to that point.
-Since we are interested in physical theories, the measure $\mu(U)$ quantifying the number of states within the region $U$.
+As we shrink $U$ to a single point $x$, we will find the probability associated to that point, and how this is done depends on the nature of $\mu$.
 
@@
 
-In the continuous case, things are a bit different.
-If $X$ is a manifold, a set of local coordinates induce a Lebesgue measure, which can be used to quantify the number of possible value combinations.
+In the continuous case, things are a bit different.\footnote{Note that ``discrete' and ``continuous'' have different meaning depending on context.
+At a set theoretic level, continuum is associated with the cardinality of a set.
+In topology, discrete and continuous spaces are characterized by  properties of open sets, not cardinality: a manifold can be given a discrete topology.
+In the context of quantifying measures in physics, discrete means that we can count elements with natural numbers, while continuous means we use real valued quantities with units, such as meters for distances, to define the size of regions.} If $X$ is a manifold, a set of local coordinates induce a Lebesgue measure, which can be used to quantify the number of possible value combinations.
 Note that a single state has measure zero, unlike the discrete case.
@@
 
-What most miss, is that the structure of classical phase space (i.e.
-symplectic structure) is exactly the structure that allows to define a quantifying measure for states that is invariant under change of observers.(TODO cite book, two phil papers) Given a change of space-time variables $q^\alpha$, conjugate momentum change covariantly, which means volumes of phase space are invariant.
-The role of quantifying measure in classical mechanics, then, is assigned to the Louisville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
-canonical coordinates).
-The amount of probability for each state is given by the Radon-Nikodym derivative $\frac{dp}{d\mu}$, which gives us the probability density.
+However, the structure of classical phase space (i.e.
+symplectic structure) is exactly the structure that allows to define a quantifying measure for states that is invariant under change of observers.(TODO cite book, two phil papers) The role of quantifying measure in classical mechanics, then, is assigned to the Louisville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
+canonical coordinates).\footnote{The volume of an infinitesimal parallelepiped is given by $dq^1 dq^2 \cdots dp_1 dp_2$.
+Since $dq^i$ is controvariant and $dp_i$ is covariant under change of reference, the volume remains the same.} The amount of probability for each state is given by the Radon-Nikodym derivative $\frac{dp}{d\mu}$, which gives us the probability density, not a the probability itself.
 
-Note that, in both cases, the quantifying measure is potentially unbounded, as the state space can have potentially infinitely many states, as it is the case for one spatial degree of freedom.
+%Note that, in both cases, the quantifying measure is potentially unbounded, as the state space can have potentially infinitely many states, as it is the case for one spatial degree of freedom.
 Therefore the quantifying measures are implicitly different from probability measures, and they require specific attention.
 
-While quantum mechanics does not obey classical probability, we do use of classical probability measure and quantifying measures in some cases.
-Measurement outcomes are described in terms of classical probability.
-If we have a discrete observable, like the energy level of a harmonic oscillator, we would use the counting measure; if we have a continuous observable, like position, we would use the Lebesgue measure associated to the variable.
-Preparations are also in terms of classical probability, and we also use both types of quantifying measures.
-For example, the maximally mixed state for a qubit can be written as $\frac{1}{2} | 0 \> \< 0 | + \frac{1}{2} | 1 \> \< 1 | $ or as a uniform mixture over all possible states $\int_\mathcal{H} \frac{1}{4\pi} | \psi \> \< \psi | d\Omega$.
+Quantum mechanics, instead, does not provide a clear state quantifying measure.
+When we characterizing preparations and measurement outcomes, we may use the discrete or continuous quantifying measures in different cases.
+For example, consider a particle.
+If we are interested in a discrete observable, like the energy level of a harmonic oscillator, we would use the counting measure; if we are interested in a continuous observable, like position, we would use the Lebesgue measure associated to the variable.
+As for preparations, the same maximally mixed state for a qubit can be written as discrete mixture $\frac{1}{2} | 0 \> \< 0 | + \frac{1}{2} | 1 \> \< 1 | $ or as a continuous uniform mixture over all possible states $\int_\mathcal{H} \frac{1}{4\pi} | \psi \> \< \psi | d\Omega$.
 
-Note how, in both cases, the quantifying measure is selected by the type of preparation or measurement, even though the state space is the same.
-But this means that these quantifying measure cannot be possibly quantifying the number of states as they would give contradicting results.
-This makes the problem rather crisp: the space of preparations and measurement outcomes are both classical, yet they cannot be connected by classical means.
+To be precise, the quantifying measures we considered are really on preparations and measurements, not the states themselves.
+It was the discrete or continuous character of the space of preparations that selected the appropriate space with the appropriate measure.
+But we can have different preparations and measurement spaces, the state space of system should always be the same, with always the same way to count states!
+This, in other words, highlights the problem of quantum probabilities: while the space of preparations and measurement outcomes are both classical, they cannot be connected by classical means.
 That is, we cannot simply put a classical probability of transition between the two because we are possibly connecting rather different quantifying measures.
@@
 To sum up, the typical foundational question is: how do we generalize probability measures for quantum mechanics?
-We are asking something different: how do we generalize measures that quantify the number of states in quantum mechanics?
+We are asking something different: given that state-quantifying measures play an equally fundamental role, how do we quantify the number of states in quantum mechanics?
 
@@
 
-While the standard structure of quantum mechanics does not have an answer to our previous question, statistical mechanics and quantum information theory have an implicit one.
-Statistical mechanics tells us that the thermodynamic entropy of a system, which is a physically meaningful quantity that has numerous empirical consequences, corresponds to the Shannon entropy for a classical discrete system, the Gibbs entropy\footnote{By Gibbs entropy we mean the differential Shannon entropy $-\int_X \rho \log \rho d\mu$ calculated over phase space using position and momentum (i.e.
+While the standard structure of quantum mechanics does not give us an answer to our question, statistical mechanics and quantum information theory have an implicit answer.
+Statistical mechanics tells us that the thermodynamic entropy of a system, which is a physically meaningful quantity that has numerous empirical consequences, corresponds to the Shannon entropy for a classical discrete system, the Gibbs entropy\footnote{By Gibbs entropy we mean the differential Shannon entropy $-\int \rho \log \rho d\mu$ calculated over phase space using position and momentum (i.e.
 canonical coordinates).} for a classical continuous system, and the von Neumann entropy for a quantum system.
@@
 Physically, the Boltzmann constant is needed to define temperature in Kelvin.
-tatistical mechanics can be equivalently formulated in terms of beta $\beta = \frac{1}{k_B T}$.TODO: cite the unit definition paper} In classical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Louisville volume in the continuous case.
+Statistical mechanics can be equivalently formulated in terms of beta $\beta = \frac{1}{k_B T}$.TODO: cite the unit definition paper} In classical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Louisville volume in the continuous case.
 The two expressions for entropy are linked as the Shannon/Gibbs entropy over a uniform distribution matches the logarithm of the measure of the support.
 That is, if $\rho_U$ is a uniform distribution over $U$, then $S(\rho_U) = \log \mu(U)$.
-
 In quantum mechanics, as we said before, we do not have a measure to quantify states.
 There is, however, a somewhat related expression: the maximum entropy attainable by a quantum system described by a Hilbert space $\mathcal{H}$ is the dimensionality $\dim_{\mathbb{C}}(\mathcal{H})$ of the space.
+Therefore, $ S(\rho_U) \leq \dim_{\mathbb{C}}(\text{span}(U))$
 
-A similar connections can be established in information theory.
+A similar connection can be established in information theory.
 The Shannon/Gibbs/von Neumann entropy quantifies the minimum number of bits required on average to send a message from a given source.\footnote{For a more general characterization of entropy that works in statistical mechanics, information theory and other contexts, and a more precise discussion of the continuous case, see TODO: cite our paper.} This, again, is a quantity that leads to empirical consequences.
 If we have $n$ bits at our disposal, we have $2^n$ possible combination, possible messages, that can be encoded.
-Therefore the logarithm of the number of possible messages corresponds to the entropy.
-Each different message can be encoded into a different state, therefore $\log \mu(U)$ tells us the number of bits that can be encoded using the states in that region.
+Therefore the logarithm of the number of available messages corresponds to the number of bits that can be encoded using those messages.
+Physically, we can use different states to encode different messages, therefore $\log \mu(U)$ tells us the number of bits that can be encoded using the states in that region.
 This will exactly correspond to the number of bits required by a source characterized by a uniform probability distribution over that number of symbols.
@@
 
-The simplest case is when $U = \{ \psi \}$ is a single state.
-In this case, we have $S(|\psi\>\<\psi|) = 0$ and therefore $\mu({\psi}) = 2^{0} = 1$.
-Note that this is the same as the classical discrete case.
+We can start by calculating the simplest case, which is when $U = \{ \psi \}$ is a single state.
+In this case, we have $S(|\psi\>\<\psi|) = 0$ and therefore $\mu(\{\psi\}) = 2^{0} = 1$.
+The measure returns one for each state as in the classical discrete case.
 
@@
 If the states are orthogonal, then $p=0$, and $S(\rho) = 1$ therefore $\mu(U) = 2^1 = 2$.
-In this case we do have that $\mu(\{ \psi, \phi \}) = \mu(\{\psi\}) + \mu(\{\phi\})$, and therefore $\mu$ is additive.
-We can also see that if $U$ is an orthonormal basis, the von Neumann entropy $S(U) = \log |U|$ is the logarithm of the cardinality of $U$.
-Therefore the quantifying measure $\mu(U) = |U|$ becomes the counting measure.
+In this case we have that $2 = \mu(\{ \psi, \phi \}) = \mu(\{\psi\}) + \mu(\{\phi\}) = 1 + 1$, and therefore $\mu$ is additive.
+Moreover, if $U$ is an orthonormal basis, the von Neumann entropy $S(U) = \dim_{\mathbb{C}}(\text{span}(U)) = \log |U|$ is the logarithm of the cardinality of $U$.
+Therefore the quantifying measure $\mu(U) = |U|$ is equal to the counting measure.
 But this is a special case that corresponds to the highest entropy reachable with a mixture.
@@
 
-To sum up, the physically motivated way to quantify the number of states in quantum mechanics is not additive. When counting states in quantum mechanics, $1+1 \leq 2$.
+To sum up, the physically motivated way to quantify the number of states in quantum mechanics is, in general, not additive. When counting states in quantum mechanics, $1+1 \leq 2$.
 
@@
 While unusual claims are common in quantum mechanics, we want to have a clear, physically tenable, reason as to why the quantifying measure for quantum states must be non-additive.
-In the context of a spin 1/2 system, for example, why is it that $|z^+\>$ and $|z^-\>$ count as two states while $|z^+\>$ and $|x^+\>$ count as less than two?
+In the context of a spin 1/2 system, for example, why is it that $\{ |z^+\>, |z^-\>\}$ count as two states while $\{ |z^+\>, |x^+\> \}$ count as less than two?
 
@@
 The selection of the measurement means a difference in physical process of either the preparation or measurement device.
-Choosing which direction of spin to prepare or measure means, at some point, orienting some polarizer or magnetic field.
+Choosing which direction of spin to prepare or measure means, at some point, orienting some polarizer, magnetic field or equivalent change to a physical device.
 Something is changed in the physical world, not just on the system, but around the system, on its environment.
-If Amanda claims that a particle is in the $|z^+\>$ state, then Boris is able to infer something about the preparation/measuring device.\footnote{This is not at all particular to quantum mechanics.
+If Amanda claims that a particle is in the $|z^+\>$ state, then Boris is able to infer something about Amanda's preparation/measuring device.\footnote{This is not at all particular to quantum mechanics.
 We can study the motion of a cannonball here on earth, but not on the surface of the sun, as the cannonball would be immediately vaporized.
@@
 Given that we have, in general, a correlation between system and environment, it is natural to understand why the contextuality leads to a non-additive measure.
-Two states, in fact, should be counted as two only if they lead to a change in the system.
+Two states, in fact, should be counted as two only if they lead to a change in the system alone.
 That is, \textbf{states should be quantified all-else-being-equal}.
@@
 
-Before trying to answer that question, let's make a direct comparison between the quantifying measures for the three cases: the counting measure $\mu_d$ for the classical discrete case, the Louisville measure $\mu_c$ for the classical continuous case and the newly defined quantum measure $\mu_d$ for the quantum case.
+Before trying to answer that question, let's make a direct comparison between the quantifying measures for the three cases: the counting measure $\mu_d$ for the classical discrete case, the Louisville measure $\mu_c$ for the classical continuous case and the newly defined quantized measure $\mu_q$ for the quantum case.
 This requires a space where all three can be defined, which fortunately exists: the surface of a sphere.
@@
 The Louisville measure returns the solid angle, which is always going to be finite as it must be less than or equal to $4\pi$.
-The von Neuman entropy of a mixed state over a two state system is at most one, and therefore the measure must be less than or equal to $2$.
+The von Neuman entropy of a mixed state over a two state system is between zero and one, and therefore the measure must be between one and $2$.
 However, given that an open set on a sphere has infinitely many point, the counting measure over a finite region will be infinite.
@@
 So, in the light of this, what is quantization?
-If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see TODO libro} If the regions are large enough, the Louisville (or Lebesgue) measure will work.
+If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see TODO libro} If the regions are large enough, the Louisville (or Lebesgue) measure will work.
 But as we shrink a region we will encounter, given the additivity, regions that contain less than one state, which does not make sense.
@@
 In our view, this provides good conceptual reasons to understand the failure of classical mechanics, the need for the departure from classical ideas and the need for quantum theory.
-We believe this is already sufficient reason to start a research program to explore these ideas fully.
-However, since most physicists are not interested in this type of ``clean-up work'', let's see this is, in our mind, critical for the development of future physical theories.
+We believe this, by itself, is already sufficient reason to start a research program to explore these ideas fully.
+However, since most physicists do not seem to be interested in this type of ``clean-up work'', we can easily argue that a full characterization of the quantized measure is critical for the development of future physical theories.
 
@@
 
-As we showed before, this line of reasoning gives us a good conceptual understanding of what quantization is: the quantifying measure for states must have a unitary lower bound.
+We have seen that quantization can be understood as putting a lower unitary bound on the state quanfication: no set of states can have fever than one state.
 If we were to quantize space-time, what is it that we are quantizing?
@@
 The most fundamental theories that are available to us are field theories.
-In a field theory, the state is described by the value of the fields at each point.
+In a field theory, the state is described by the value of the fields at each point in space.
 At each point, the value of the field is independent from the others: the degrees of freedom form a continuum.\footnote{If we restrict ourselves to continuous fields, then we only need to specify the value of the fields on a countable dense subset (i.e.
@@
 
-We have again the same problem as before: measures on the continuum are not invariant under generic coordinate transformations.
-We do have a sense, however, that if we double the spatial volume, we double the number of degrees of freedom, and spatial volume is an invariant.
-Therefore we can use a measure of volume constructed from the metric tensor.
-This would be the equivalent of the Louisville measure we got from phase space, and we end up with an equivalent problem.
+If we have a region of space and we double the volume, we can imagine we are also doubling the number of degrees of freedom.
+Therefore, we can construct a volume measure from the metric tensor, which would give us an invariant volume that quantifies degrees of freedom, much in the same way that the Louisville measure quantified states.
 
@@
 This does not make sense.
-If we want to quantize space, then, we need to put a unitary lower bound to the measure.
-But now our measure cannot be additive: what does it mean that two degrees of freedom count less than two?
+As before, we need to put a unitary lower bound to the measure and our measure cannot be additive.
+We need a quantized measure.
+Quantizing space-time, means putting a quantized measure on the number of degrees of freedom.
 
-Before answering that question, let's add time.
+What does it mean that two degrees of freedom count less than two?
+Before answering that question, let's consider time.
 We said that if we double the spatial volume, we double the number of degrees of freedom.
@@
 
-So, what does it mean that two degrees of freedom count less than two?
-It could simply mean that they are not independent.
-That is, the values of the field at two distant points can be chosen arbitrarily, but as the points get sufficiently close, this would not be the case.
-This would make sense.
+The same field at two points, then, may count less than to degrees of freedom simply because they are not independent.
+The values of the field at two distant points can be chosen arbitrarily, therefore the measure would be additive at long range.
+But as the points get sufficiently close, this would not be the case, so at short scale we need a quantized measure to quantify independent degrees of freedom.
 Note that the lack of independence of the field values in close-by regions may mean that high frequencies of the fields are damped: the values have to become closer.
-The quantized quantifying measure of the degrees of freedom would effectively provide a cutoff.
-This may provide a natural high energy cutoff for quantum field theories.
+A quantized measure of the degrees of freedom would effectively provide a cutoff.
+Therefore we have a story that is very well motivated and provide a possible solution for high energy divergences in quantum field theories.
 
-In our view, the arguments laid here highlight an undeniable problem and seem to offer a natural step to take in line with the change from classical to quantum theory.
-While the development of a quantum theory of gravity is not our primary interest, the mathematical tools needed to address the above problem are of the same nature as the one we propose to develop a non-additive measure theory to quantify quantum states.
+While the development of a quantum theory of gravity is not our primary interest, the mathematical tools required to characterize the non-additive quantized measure seem to be of similar nature as the ones that may be used to quantize space-time.
 
@@
 
-Having outlined the scope and motivation for this new approach, we now outline the main questions we think need to be addressed.
+Having outlined the scope and motivation for this new approach, we now outline the main questions we think need to be addressed by this new research program.
 
@@
 For the counting measure, once we define the value for a single point, typically one, the whole measure is determined.
-Similarly, once we have defined a real variable, the measure is determined.
+Similarly, once we have defined a real variable, the Lebesgue measure is determined.
 The Louisville requires a bit more structure, but once the canonical coordinates are chosen, the measure is determined as well.
@@
 
-While we have found some literature on non-additive measures and set functions (CITE), together with modified theories of integration, these seem to concentrate on monotonic measure, that is measure that increase as the set increase.
+While we have found some literature on non-additive measures and set functions (CITE), together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increase.
 Unfortunately, the quantized measure does not have this feature.
@@
 The set is bigger, but the measure is smaller.
-Therefore we may need to develop an appropriate theory.
+Therefore there may be new mathematical ground to be explored.
 
-To achieve the previous point, we will need to understand the necessary conditions for the quantifying measure.
+We may want to understand the necessary conditions for a quantifying measure.
 Some may be derived from the strict concavity of entropy, given the connection between the two.
 Furthermore, note that the relationship between the measure over two points, the entropy of the equal mixture of two states and the probability of transition between two states are invertible.
-Therefore the relationship between two states defines the geometry of the whole state space.
+This means that the relationship between two states defines the geometry of the whole state space.
 Therefore there must be further consistency requirements that can be identified.
 
-Another thing that need to be understand is how the quantized measure behaves with composite systems.
+Another issue that needs to be understood is how the quantized measure behaves in composite systems and across multiple degrees of freedom.
 In the classical case, the measure is defined on the product of the $\sigma$-algebra.
@@
 However, we have found a connection between the two: if the symplectic form is expressed in position and velocity, instead of canonical coordinate, the metric tensor appears as the position-velocity component.
-It then can be understood as defining volumes in position-velocity instead in position-momentum.
+The metric tensor can then be understood as defining volumes in position-velocity instead in position-momentum.
 This may allow us to transfer the quantized measure on the tangent bundle and then, hopefully, to the space itself.
 
-There is therefore much work to be done, but one thing we want to stress: a clear and precise physical idea is driving the development of possibly new math.
-We are not looking for mathematical extensions and trying them to see whether something useful can be achieved.
-We have a goal, quantification of states .
-
- are looking to quantify states  We are not introducing something like negative probability, we have a clear understanding of what the physical significance of the generalized measure is.
-The goal is the quantification of states in the contextual case.
+Finally, we want to stress that, contrary to what happens in other programs, the physical motivation for the mathematical generalization are tightly driven by the physics.
+We are not introducing concepts, like negative probability, have no clear direct physical meaning.
+The goal is the quantification of states.
 

### 2023-10-23 BareMinima/SetTheory/BareMinimumSetTheory.tex
BM-SetTheory: minor correction
SetTheory: using "connected" instead of "total" for relationships

@@
 		\item[Transitive.] If $a$ is related to $b$ and $b$ is related to $c$, then $a$ is related to $c$. That is, if $aRb$ and $bRc$ then $aRc$.
-		\item[Total.] Given two distinct elements, one is related to the other.That is, given $a \neq b$, either $aRb$ or $bRa$. (This has no relation to left/right-total)
+		\item[Connected.] Given two distinct elements, one is related to the other. That is, given $a \neq b$, either $aRb$ or $bRa$.
 	\end{description}
@@
 	A \textbf{partial order} \marginleft{Order, equivalence: $\leq$, $<$, $\equiv$} (noted for example $\leq$, $\lesssim$, $\preceq$)  is a homogeneous binary relationship that is reflexive, antisymmetric and transitive.
-A \textbf{linear order} or \textbf{total order} is an order that is total.
+A \textbf{linear order} or \textbf{total order} is an order that is connected.
 An order is \textbf{strict} (noted for example $<$, $\prec$) if it is irreflexive instead of reflexive.
+A \textbf{strict linear order} or \textbf{strict total order} is an order that is strict and connected.
 
@@
 		\item Trichotomy: given two sets, either they have equal cardinality, or one has lower cardinality than the other (i.e. $|A| < |B|$, $|A| = |B|$ or $|A| > |B|$)
-		\item Zorn's lemma: every non-empty poset (i.e. partially ordered set) for which every chain (i.e. totally ordered subset) has an upper bound contains at least one maximal element
+		\item Zorn's lemma: every non-empty poset (i.e. partially ordered set) for which every chain (i.e. linearly ordered subset) has an upper bound contains at least one maximal element
 		\item Every vector space has a basis

### 2023-10-24 Proceedings/2023-QIP/QuantumMeasure.tex
QIP: reviewed with Christine

@@
 
-As part of our overall project Assumptions of Physics, we have developed an approach, called reverse physics, which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive the laws.
+As part of our overall project Assumptions of Physics, we have developed an approach, called reverse physics, which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive them.
 We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.
 The former (i.e.
-the symplectic structure) is the only structure that allows to quantify states and configurations over independent degrees of freedom in a way that is observer independent.
+the symplectic structure) is the only structure that allows one to quantify states and configurations over independent degrees of freedom in a way that is frame independent.
 If one imposes deterministic and reversible evolution, state quantification must be preserved, leading to Hamiltonian evolution (i.e.
@@
 Given the formal similarity between classical and quantum mechanics, we believe a similar approach is possible for the latter.
-Which raises the question: what is the quantum analogue of the Louisville measure, the measure that quantifies states in the classical theory?
+This raises the question: what is the quantum analogue of the Liouville measure, the measure that quantifies states in the classical theory?
 
@@
 
-The goal of this article is to propose and articulate the goals of this new line of research.
+The aim of this article is to propose and articulate the goals of this new line of research.
 It provides the overall picture in which future works will add further mathematical detail and physical insights.
 In section II we will show how measure theory is used to quantify elements in a set and how this is needed to assign probability and probability densities to each element.
-In section III we will see how statistical mechanics and information theory mandate a link between entropy and state quantification, which we will use to define the quantum analogue of the Louisville measure.
+In section III we will see how statistical mechanics and information theory mandate a link between entropy and state quantification, which we will use to define the quantum analogue of the Liouville measure.
 We will find that this measure is not additive.\footnote{Technically, measures are defined to be additive, therefore what we find is not a measure, but a set function.
 However, this term is not widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).} In section IV we will see that this non-additivity is linked to non-contextuality.
-In section V we will see that quantization means putting a lower unitary bound on the quantification of states, while keeping a finite value over finite regions.
+In section V we will see that quantization means putting a unitary lower bound on the quantification of states, while keeping a finite value over finite regions.
 We will see that this very directly requires non-additivity and therefore non-contextuality.
 In section VI we will argue that quantization of space-time means using a quantized measure on the quantification of degrees of freedom.
-Lastly, in section VII we will outline a series of goals for a research program that aims to fully characterized the quantized measure.
+Lastly, in section VII we will outline a series of goals for a research program that aims to fully characterize the quantized measure.
 
@@
 As we said in the introduction, measure theory is not only used to characterize Kolmogorovian probability.
-More in general, measures are used to define the size of sets.
-Even in probability theory, since probability measures assigns probability to sets of point, to associate probability to points one needs two measures: the \textbf{probability measure} $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$.\footnote{Measure assigns a value to elements to a $\sigma$-algebra that can be thought as a set of experimental procedures.} Since we are interested in physical theories, the measure $\mu$ quantifies the number of states.
-Given a set $U$, the ratio $\frac{p(U)}{\mu(U)}$ given us the average probability per unit size.
+More generally, measures are used to define the size of sets.
+Even in probability theory, since probability measures assign probability to sets of points, in order to associate probability to individual points one needs two measures: the \textbf{probability measure} $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$.\footnote{Measures assign values to elements of a $\sigma$-algebra, which can be thought of as a set of experimental procedures.} Since we are interested in physical theories, the measure $\mu$ quantifies the number of states.
+Given a set $U$, the ratio $\frac{p(U)}{\mu(U)}$ gives us the average probability per unit of states.
 As we shrink $U$ to a single point $x$, we will find the probability associated to that point, and how this is done depends on the nature of $\mu$.
@@
 Therefore for a single point we simply have $\frac{p(\{x\})}{\mu(\{x\})} = \frac{p(\{x\})}{1} = p(x)$, which corresponds to the probability associated to the state $x$.
-The counting measure is fully determined by two properties: countable additivity, a defining property of measures, and that each states counts as one.
+The counting measure is fully determined by two properties: countable additivity, a defining property of measures, and that each state counts as one.
 
-In the continuous case, things are a bit different.\footnote{Note that ``discrete' and ``continuous'' have different meaning depending on context.
-At a set theoretic level, continuum is associated with the cardinality of a set.
+In the continuous case, things are a bit different.\footnote{Note that ``discrete' and ``continuous'' have different meanings depending on context.
+At a set theoretic level, ``continuum'' is associated with the cardinality of a set.
 In topology, discrete and continuous spaces are characterized by  properties of open sets, not cardinality: a manifold can be given a discrete topology.
@@
 Note that a single state has measure zero, unlike the discrete case.
-The Lebesgue measure is defined by two properties: countable additivity and that the size of parallelepipeds corresponds to the volume expressed by the product of the difference of coordinates.
-Naturally, this is a problem in physics given that the quantification of the number of states should be the same for all observers.
+The Lebesgue measure is defined by two properties: countable additivity and that the size of parallelepipeds corresponds to the volume expressed by the product of the differences of coordinates (i.e.
+$\mu(U) = \Delta x^1 \Delta x^2 \cdots \Delta x^n$).
+Naturally, this is a problem in physics given that the quantification of the number of states should be the same in all frames.
 
 However, the structure of classical phase space (i.e.
-symplectic structure) is exactly the structure that allows to define a quantifying measure for states that is invariant under change of observers.(TODO cite book, two phil papers) The role of quantifying measure in classical mechanics, then, is assigned to the Louisville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
-canonical coordinates).\footnote{The volume of an infinitesimal parallelepiped is given by $dq^1 dq^2 \cdots dp_1 dp_2$.
-Since $dq^i$ is controvariant and $dp_i$ is covariant under change of reference, the volume remains the same.} The amount of probability for each state is given by the Radon-Nikodym derivative $\frac{dp}{d\mu}$, which gives us the probability density, not a the probability itself.
+symplectic structure) is exactly the structure that allows one to define a quantifying measure for states that is invariant under change of frame.(TODO cite book, two phil papers) The role of the quantifying measure in classical mechanics, then, is assigned to the Liouville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
+canonical coordinates).\footnote{The volume of an infinitesimal parallelepiped is given by $dq^1 dq^2 \cdots dp_1 dp_2 \cdots$.
+Since $dq^i$ is contravariant and $dp_i$ is covariant under change of frame, the volume remains the same.} As we shrink the size of the set, $\frac{p(U)}{\mu(U)}$ becomes the Radon-Nikodym derivative $\frac{dp}{d\mu}$, which gives us the probability density, not the probability itself.
 
@@
 Quantum mechanics, instead, does not provide a clear state quantifying measure.
-When we characterizing preparations and measurement outcomes, we may use the discrete or continuous quantifying measures in different cases.
+When we characterize preparations and measurement outcomes, we may use the discrete or continuous quantifying measures in different cases for the same system.
 For example, consider a particle.
 If we are interested in a discrete observable, like the energy level of a harmonic oscillator, we would use the counting measure; if we are interested in a continuous observable, like position, we would use the Lebesgue measure associated to the variable.
-As for preparations, the same maximally mixed state for a qubit can be written as discrete mixture $\frac{1}{2} | 0 \> \< 0 | + \frac{1}{2} | 1 \> \< 1 | $ or as a continuous uniform mixture over all possible states $\int_\mathcal{H} \frac{1}{4\pi} | \psi \> \< \psi | d\Omega$.
+As for preparations, the same maximally mixed state for a qubit can be written as a discrete mixture $\frac{1}{2} | 0 \> \< 0 | + \frac{1}{2} | 1 \> \< 1 | $ or as a continuous uniform mixture over all possible states $\int \frac{1}{4\pi} | \psi \> \< \psi | d\Omega$.
 
@@
 It was the discrete or continuous character of the space of preparations that selected the appropriate space with the appropriate measure.
-But we can have different preparations and measurement spaces, the state space of system should always be the same, with always the same way to count states!
-This, in other words, highlights the problem of quantum probabilities: while the space of preparations and measurement outcomes are both classical, they cannot be connected by classical means.
-That is, we cannot simply put a classical probability of transition between the two because we are possibly connecting rather different quantifying measures.
+But even if we have different preparation and measurement spaces, the state space of the system should always be the same, with a single way to count states.
+This, in other words, highlights the problem of quantum probabilities: while the spaces of preparations and measurement outcomes are both classical, they cannot be connected by classical means.
+That is, we cannot simply put a classical probability of transition between the two because we would be connecting qualitatively different quantifying measures.
 
@@
 
-There is another expression for entropy, often referred to as the fundamental postulate of statistical mechanics, which is the logarithm of the count of microstates: $\log \mu(U)$.\footnote{Note that the Boltzmann constant is left out, as we will be always measured in bits or nats.
+There is another expression for entropy, often referred to as the fundamental postulate of statistical mechanics, which is the logarithm of the count of microstates: $\log \mu(U)$.\footnote{Note that the Boltzmann constant is left out, as entropy will always be expressed in bits or nats.
 Physically, the Boltzmann constant is needed to define temperature in Kelvin.
-Statistical mechanics can be equivalently formulated in terms of beta $\beta = \frac{1}{k_B T}$.TODO: cite the unit definition paper} In classical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Louisville volume in the continuous case.
+Statistical mechanics can be equivalently formulated in terms of $\beta = \frac{1}{k_B T}$.TODO: cite the unit definition paper} In classical statistical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Liouville volume in the continuous case.
 The two expressions for entropy are linked as the Shannon/Gibbs entropy over a uniform distribution matches the logarithm of the measure of the support.
@@
 In quantum mechanics, as we said before, we do not have a measure to quantify states.
-There is, however, a somewhat related expression: the maximum entropy attainable by a quantum system described by a Hilbert space $\mathcal{H}$ is the dimensionality $\dim_{\mathbb{C}}(\mathcal{H})$ of the space.
-Therefore, $ S(\rho_U) \leq \dim_{\mathbb{C}}(\text{span}(U))$
+There is, however, a somewhat related expression: the maximum entropy attainable by a quantum system corresponding to a Hilbert space $\mathcal{H}$ is the logarithm of the dimensionality $\dim_{\mathbb{C}}(\mathcal{H})$ of the space.
+Therefore, $ S(\rho_U) \leq \log (\dim_{\mathbb{C}}(\text{span}(U)))$.
 
@@
 The Shannon/Gibbs/von Neumann entropy quantifies the minimum number of bits required on average to send a message from a given source.\footnote{For a more general characterization of entropy that works in statistical mechanics, information theory and other contexts, and a more precise discussion of the continuous case, see TODO: cite our paper.} This, again, is a quantity that leads to empirical consequences.
-If we have $n$ bits at our disposal, we have $2^n$ possible combination, possible messages, that can be encoded.
-Therefore the logarithm of the number of available messages corresponds to the number of bits that can be encoded using those messages.
-Physically, we can use different states to encode different messages, therefore $\log \mu(U)$ tells us the number of bits that can be encoded using the states in that region.
+If we have $n$ bits at our disposal, we have $2^n$ possible combinations, possible messages, that can be encoded.
+Therefore the logarithm of the number of potential messages corresponds to the number of bits that can encode those messages.
+Physically, we can associate a different state to each message to be encoded.
+Therefore $\log \mu(U)$ corresponds to the number of bits.
 This will exactly correspond to the number of bits required by a source characterized by a uniform probability distribution over that number of symbols.
@@
 In this case we have that $2 = \mu(\{ \psi, \phi \}) = \mu(\{\psi\}) + \mu(\{\phi\}) = 1 + 1$, and therefore $\mu$ is additive.
-Moreover, if $U$ is an orthonormal basis, the von Neumann entropy $S(U) = \dim_{\mathbb{C}}(\text{span}(U)) = \log |U|$ is the logarithm of the cardinality of $U$.
+Moreover, if $U$ is an orthonormal basis, the von Neumann entropy $S(U) = \log(\dim_{\mathbb{C}}(\text{span}(U))) = \log |U|$ is the logarithm of the cardinality of $U$.
 Therefore the quantifying measure $\mu(U) = |U|$ is equal to the counting measure.
@@
 
-Since we saw that the quantifying measure is indeed additive for a set of orthogonal states, this should make us think about contextuality: the impossibility to assign outcomes to all potential measurement of a quantum system.
+Since we saw that the quantifying measure is indeed additive for a set of orthogonal states, this should make us think about contextuality: the impossibility to assign outcomes to all potential measurements of a quantum system.
 We have, in fact, derived that it is impossible to define an additive quantifying measure on all potential states, but we can do so on the outcomes of one measurement.
 The ability/inability to define additive measures is exactly the difference between classical/non-classical probability.
-We are definitely onto something.
 
 While contextuality is often presented as a somewhat abstract, almost metaphysical, property of quantum mechanics, we have a rather more down-to-earth view.
-The selection of the measurement means a difference in physical process of either the preparation or measurement device.
-Choosing which direction of spin to prepare or measure means, at some point, orienting some polarizer, magnetic field or equivalent change to a physical device.
-Something is changed in the physical world, not just on the system, but around the system, on its environment.
+The choice of the measurement means a difference in physical process of either the preparation or measurement device.
+Choosing which direction of spin to prepare or measure means, at some point, orienting some polarizer, magnetic field or making an equivalent change to a physical device.
+Something is changed in the physical world, not just in the system, but around the system, in its environment.
 If Amanda claims that a particle is in the $|z^+\>$ state, then Boris is able to infer something about Amanda's preparation/measuring device.\footnote{This is not at all particular to quantum mechanics.
 We can study the motion of a cannonball here on earth, but not on the surface of the sun, as the cannonball would be immediately vaporized.
-The mere claim that a system exists implies that the boundary conditions necessary for that system to exist are satisfies.
-In this sense, all physics is contextual.} Therefore we can understand the context simply as those conditions on the environment that are necessary on the system to be found in that particular state.
+The mere claim that a system exists implies that the boundary conditions necessary for that system to exist are satisfied.
+In this sense, all physics is contextual.} Therefore we can understand the context simply as those conditions on the environment that are necessary for the system to be found in that particular state.
 A set of orthogonal states, then, has the property that they can be found in the same external conditions.
 
-Given that we have, in general, a correlation between system and environment, it is natural to understand why the contextuality leads to a non-additive measure.
+Given that we have, in general, a correlation between system and environment, it is natural to understand how contextuality leads to a non-additive measure.
 Two states, in fact, should be counted as two only if they lead to a change in the system alone.
@@
 While two states from different contexts are indeed two distinct states, stating we transitioned from one to the other does not give us information only about the system: it tells us something about the environment as well.
-The degree of incompatibility of observables, then, ultimately measure the degree of change when realizing the context physically.
+The degree of incompatibility of observables, then, ultimately measures the degree of change when realizing the context physically.
 
@@
 
-Before trying to answer that question, let's make a direct comparison between the quantifying measures for the three cases: the counting measure $\mu_d$ for the classical discrete case, the Louisville measure $\mu_c$ for the classical continuous case and the newly defined quantized measure $\mu_q$ for the quantum case.
+Before trying to answer that question, let's make a direct comparison between the quantifying measures for the three cases: the counting measure $\mu_d$ for the classical discrete case, the Liouville measure $\mu_c$ for the classical continuous case and the newly defined quantized measure $\mu_q$ for the quantum case.
 This requires a space where all three can be defined, which fortunately exists: the surface of a sphere.
@@
 The counting measure returns one, and so does the quantum measure.
-The Louisville measure, however, returns zero.
+The Liouville measure, however, returns zero.
 Second, we want to compare the behavior on a finite region, an open set.
-The Louisville measure returns the solid angle, which is always going to be finite as it must be less than or equal to $4\pi$.
-The von Neuman entropy of a mixed state over a two state system is between zero and one, and therefore the measure must be between one and $2$.
-However, given that an open set on a sphere has infinitely many point, the counting measure over a finite region will be infinite.
+The Liouville measure returns the solid angle, which is always going to be finite as it must be less than or equal to $4\pi$.
+The von Neumann entropy of a mixed state over a two-state system is between zero and one, and therefore the measure must be between one and two.
+However, given that an open set on a sphere has infinitely many points, the counting measure over a finite region will be infinite.
 Lastly, we want to compare the additivity.
@@
 	\item single states count as one
-	\item finite continuous regions correspond to finite information (and therefore finite state quantification)
+	\item finite continuous regions correspond to finite state quantification
 	\item state quantification is additive on disjoint regions.
@@
 These seem very reasonable properties to assume for a quantifying measure of states.
-\textbf{Unfortunately, these three conditions are incompatible}.
+\textbf{However, these three conditions are incompatible}.
 If each state counts as one, and the measure is additive, then a finite continuous region must be infinite as it contains infinite points.
@@
 The counting measure rejects the second property.
-The Louisville (and Lebesgue) measure discards the first.
+The Liouville (and Lebesgue) measure discards the first.
 The quantum measure sacrifices the third.
 
-So, in the light of this, what is quantization?
-If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see TODO libro} If the regions are large enough, the Louisville (or Lebesgue) measure will work.
+So, in light of this, what is quantization?
+If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see TODO libro} If the regions are large enough, the Liouville (or Lebesgue) measure will work.
 But as we shrink a region we will encounter, given the additivity, regions that contain less than one state, which does not make sense.
-Low value of the quantifying measure correspond to low values of entropy and measure one corresponds do zero entropy.
-Regions with measure less than one correspond to negative thermodynamic entropy, which also does not make sense.
-In fact, classical statistical mechanics does fail in those regimes: the classical equations of state for an ideal gas to predict negative entropy, which is fixed by the Fermi and Bose statistics.
+Note that measure less than one corresponds to negative thermodynamic entropy, which does not make sense either.
 
-Quantization, then, is fixing the counting measure so that it is bounded from below by one (and the entropy by zero): each state counts as one.
+Quantization, then, is fixing the quantifying measure so that it is bounded from below by one (and the entropy by zero): each state counts as one.
 However, this comes at a price: non-additivity.
-This non-addivity is ultimately why classical probability has to fail as well, given that additivity is one of the axioms.
-It is also the same reason as why quantization implies contextuality: the non-addivity means that some states cannot be prepared at-all-else-being equal.
-%(https://en.wikipedia.org/wiki/Sackur%E2%80%93Tetrode_equation)
+This non-additivity is ultimately why classical probability has to fail as well, given that additivity is one of its axioms.
+It is also the same reason as why quantization implies contextuality: the non-additivity means that some states cannot be prepared at-all-else-being-equal.
 
-In our view, this provides good conceptual reasons to understand the failure of classical mechanics, the need for the departure from classical ideas and the need for quantum theory.
+In our view, this provides a good conceptual basis to understand the failure of classical mechanics, the need for the departure from classical ideas and the need for quantum theory.
 We believe this, by itself, is already sufficient reason to start a research program to explore these ideas fully.
@@
 
-We have seen that quantization can be understood as putting a lower unitary bound on the state quanfication: no set of states can have fever than one state.
+We have seen that quantization can be understood as putting a unitary lower bound on the state quantification: no set of states can have fewer than one state.
 If we were to quantize space-time, what is it that we are quantizing?
@@
 rational values of the coordinates).
-Though these kind of details do not change the discussion: a Lebesgue measure on the rationals have the same characteristics of the one extended to the reals.} We can therefore pose the question: how do we quantify the degrees of freedom?
+However, these details do not change the discussion: a Lebesgue measure on the rationals has the same characteristics as the one extended to the reals.} We can therefore pose the question: how do we quantify the degrees of freedom?
 
 If we have a region of space and we double the volume, we can imagine we are also doubling the number of degrees of freedom.
-Therefore, we can construct a volume measure from the metric tensor, which would give us an invariant volume that quantifies degrees of freedom, much in the same way that the Louisville measure quantified states.
+Therefore, we can construct a volume measure from the metric tensor, which would give us an invariant volume that quantifies degrees of freedom, much in the same way that the Liouville measure quantified states.
 
@@
 This does not make sense.
-As before, we need to put a unitary lower bound to the measure and our measure cannot be additive.
+As before, we need to put a unitary lower bound on the measure and our measure cannot be additive.
 We need a quantized measure.
-Quantizing space-time, means putting a quantized measure on the number of degrees of freedom.
+Quantizing space-time, then, means putting a quantized measure on the number of degrees of freedom.
 
@@
 Suppose we have a space-time region and we double the time, are we doubling the number of degrees of freedom?
-Well, we certainly doubling the support of the function we have to specify for the field configuration, but we wouldn't consider that doubling the degrees of freedom.
-Intuitively, we just think of those as being the same degrees of freedom that are getting different values.
+Well, we are certainly doubling the support of the function we have to specify for the field configuration, but we wouldn't consider that doubling the degrees of freedom.
+Intuitively, we just think of those as being the same degrees of freedom that are getting different values in time.
 This is because we expect the future and past values to be linked by some dynamical equations, so the future values would be linked, in some way, to the past values.
 That is, the past and future value of the same field at some point are not independent.
+This may give us an insight into the physical meaning of the non-additivity of quantification of the number of degrees of freedom.
 
-The same field at two points, then, may count less than to degrees of freedom simply because they are not independent.
+The same field at two spatial points may count less than two degrees of freedom simply because their values are not independent.
 The values of the field at two distant points can be chosen arbitrarily, therefore the measure would be additive at long range.
@@
 A quantized measure of the degrees of freedom would effectively provide a cutoff.
-Therefore we have a story that is very well motivated and provide a possible solution for high energy divergences in quantum field theories.
+Therefore we have a story that is very well motivated and provides a possible solution for high energy divergences in quantum field theories.
 
@@
 
-Both the counting measure and the Lebesgue/Louisville measures are unique in some sense.
+Both the counting measure and the Lebesgue/Liouville measures are unique in some sense.
 For the counting measure, once we define the value for a single point, typically one, the whole measure is determined.
 Similarly, once we have defined a real variable, the Lebesgue measure is determined.
-The Louisville requires a bit more structure, but once the canonical coordinates are chosen, the measure is determined as well.
-Therefore the first question is whether these quantized measure are also unique in some sense.
+The Liouville measure requires a bit more structure, but once the canonical coordinates are chosen, the measure is determined as well.
+Therefore the first question is whether these quantized measures are also unique in some sense.
 
-While we have found some literature on non-additive measures and set functions (CITE), together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increase.
+While we have found some literature on non-additive measures and set functions (CITE), together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increases.
 Unfortunately, the quantized measure does not have this feature.
-Consider the set $\{ z^+, z^- \}$.
+Consider the set $\{ |z^+\>, |z^-\> \}$.
 An equal mixture will return the maximally mixed state which, as we said, would correspond to a measure equal to 2.
-Now consider the set $\{ z^+, z^-, x^+ \}$.
-This will not correspond to the maximally mixed state, and therefore the measure will be less than two.
+Now consider the set $\{ |z^+\>, |z^-\>, |x^+\> \}$.
+An equal mixture of them will not correspond to the maximally mixed state, and therefore the measure will be less than two.
 The set is bigger, but the measure is smaller.
@@
 Some may be derived from the strict concavity of entropy, given the connection between the two.
-Furthermore, note that the relationship between the measure over two points, the entropy of the equal mixture of two states and the probability of transition between two states are invertible.
+Furthermore, note that the pairwise relationships between the measure over two points, the entropy of the equal mixture of two states and the probability of transition between two states are all invertible.
 This means that the relationship between two states defines the geometry of the whole state space.
@@
 Another issue that needs to be understood is how the quantized measure behaves in composite systems and across multiple degrees of freedom.
-In the classical case, the measure is defined on the product of the $\sigma$-algebra.
-Ideally, we would like to show that this simple procedure cannot support a quantized measure, and the $\sigma$-algebra of tensor product must be taken.\footnote{Technically it would be the $\sigma$-algebra projective space coming from the tensor product of the Hilbert spaces associated with the individual systems.}
+In the classical case, the measure is defined on the product of the corresponding $\sigma$-algebras.
+Ideally, we would like to show that this simple procedure cannot support a quantized measure, and the $\sigma$-algebra of the tensor product must be taken.\footnote{Technically it would be the $\sigma$-algebra projective space coming from the tensor product of the Hilbert spaces associated with the individual systems.}
 
-The first spaces we should study are the symplectic ones as these are the ones that allow for observer independent densities.
-If we want to extend to a quantization of space-time, however, Riemannian spaces need to be studies as well.
-However, we have found a connection between the two: if the symplectic form is expressed in position and velocity, instead of canonical coordinate, the metric tensor appears as the position-velocity component.
+The first spaces we should study are the symplectic ones as these are the ones that allow for frame independent densities.
+If we want to extend to a quantization of space-time, however, Riemannian spaces need to be studied as well.
+However, we have found a connection between the two: if the symplectic form is expressed in position and velocity, instead of canonical coordinates, its position-velocity component corresponds to the metric tensor.
 The metric tensor can then be understood as defining volumes in position-velocity instead in position-momentum.
-This may allow us to transfer the quantized measure on the tangent bundle and then, hopefully, to the space itself.
+This may allow us to transfer the quantized measure onto the tangent bundle and then, hopefully, to the space itself.
 
-Finally, we want to stress that, contrary to what happens in other programs, the physical motivation for the mathematical generalization are tightly driven by the physics.
-We are not introducing concepts, like negative probability, have no clear direct physical meaning.
+Finally, we want to stress that, contrary to what happens in other programs, the physical motivations for the mathematical generalization are tightly driven by the physics.
+We are not introducing concepts that, like negative probability, have no clear direct physical meaning.
 The goal is the quantification of states.
@@
 We have seen that, in physics, measure theory is not only used to characterize classical probability but also to quantify states in state spaces.
-Ideally, such measure would have the following three conditions:
+Ideally, such a measure would have the following three conditions:
 \begin{enumerate}
 	\item single states count as one
-	\item finite continuous regions correspond to finite information (and therefore finite state quantification)
+	\item finite continuous regions correspond to finite state quantification
 	\item state quantification is additive on disjoint regions.
@@
 However, these conditions are not compatible, and only two of the three can be chosen.
-The classical discrete cases discards the second; the classical continuous case discards the first; the quantum case discards the third.
-Quantization, then, can be understood as putting a unitary lower bound to the measure that quantifies states.
+The classical discrete case discards the second; the classical continuous case discards the first; the quantum case discards the third.
+Quantization, then, can be understood as putting a unitary lower bound on the measure that quantifies states.
 
 The lack of additivity on the whole space makes it impossible to use classical probability on the state space as a whole.
-Physically, the issue is that states must be counted at-all-else-being-equal, feature broken by contextuality.
+Physically, the issue is that states must be counted at-all-else-being-equal, a feature broken by contextuality.
 But if the states do belong to the same context (e.g.
@@
 This tells us very directly why quantization requires contextuality, and why classical probability works within the same context.
-The fact that single states and and finite continuous regions have both finite measure also tells us why quantum mechanics mixes features of the classical and discrete cases.
+The fact that single states and finite continuous regions both have finite measure also tells us why quantum mechanics mixes features of the classical discrete and continuous cases.
 
-We believe that further study and characterization of the non-additive quantized measure will not only bring more clarity in the interplay between the above concept, but it may be necessary for the development of future physical theory.
-Given that, in a field theory, the number of independent degrees of freedom is taken to be proportional to the spatial volume, quantization of space time can be understood as the quantization of the measure that quantifies the indipendent degrees of freedom.
+We believe that further study and characterization of the non-additive quantized measure will not only bring more clarity to the interplay among the above concepts, but it may be necessary for the development of future physical theories.
+Given that, in a field theory, the number of independent degrees of freedom is taken to be proportional to the spatial volume, quantization of space-time can be understood as the quantization of the measure that quantifies the independent degrees of freedom.
 The development of physically motivated mathematical tools may help in that regard.

### 2023-10-25 Proceedings/2023-QIP/QuantumMeasure.tex
QIP: added abstract, citations and acknowledgements

@@
 \begin{abstract}
+	Measure theory is used in physics, not just to capture classical probability, but also to quantify the number of states.
+In previous works, in fact, we found that state quantification plays a foundational role in classical mechanics, and, therefore, we set ourselves to construct the quantum equivalent of the Liouville measure.
+Unlike the classical counterpart, this quantized measure is non-additive and has a unitary lower bound (i.e.
+no set of states cannot have less than one state).
+Conversely, requiring that state quantification is finite for finite continuous regions and unitary for each state already implies non-additivity, which implies the failure of classical theory.
+In this article we show these preliminary results and outline a new line of inquiry that may provide a different insights into the foundations of quantum theory.
+Additionally, this new approach may turn out to be useful to those interested in a quantized theory of space-time, as we believe this requires a quantized measure for the quantification of the independent degrees of freedom.
 \end{abstract}
@@
 non-Kolmogorov) probability theory.
-To obviate this problem, various alternatives and extensions to measure theory and probability theory have been developed with different goals.
-Our interest in measure theory, however, is not in its use to capture classical probability.
+To obviate this problem, various alternatives and extensions to measure theory and probability theory have been developed with different goals.\cite{groenewold1946principles, moyal1949quantum, gleason1975measures, sorkin1994quantum, hamhalter2013quantum, svozil2022extending, monchietti2023measure} Our interest in measure theory, however, is not in its use to capture classical probability.
 Measures play another crucial use in physical theories: ``counting'', or better ``quantifying'', states.
@@
 
-As part of our overall project Assumptions of Physics, we have developed an approach, called reverse physics, which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive them.
-We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.
-The former (i.e.
+As part of our overall project Assumptions of Physics,\cite{aop-book} we have developed an approach, called reverse physics,\cite{aop-phys-ReversePhysics} which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive them.
+We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.\cite{aop-phil-HamiltonianPrivilege,aop-phil-Hamiltonianinformation} The former (i.e.
 the symplectic structure) is the only structure that allows one to quantify states and configurations over independent degrees of freedom in a way that is frame independent.
@@
 a symplectomorphism).
-All elements of classical mechanics, including canonical variables, the Lagrangian and the action principle, can be rederived and understood in terms of quantifying the flow of states.
+All elements of classical mechanics, including canonical variables, the Lagrangian and the action principle, \cite{aop-phys-ActionPrincipleInterpretation} can be rederived and understood in terms of quantifying the flow of states.
 Given the formal similarity between classical and quantum mechanics, we believe a similar approach is possible for the latter.
@@
 We will find that this measure is not additive.\footnote{Technically, measures are defined to be additive, therefore what we find is not a measure, but a set function.
-However, this term is not widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature (TODO: cite a couple of books on non-additive measure theory).} In section IV we will see that this non-additivity is linked to non-contextuality.
+However, this term is not widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature.
+\cite{nonadditive2016, denneberg1994non}} In section IV we will see that this non-additivity is linked to non-contextuality.
 In section V we will see that quantization means putting a unitary lower bound on the quantification of states, while keeping a finite value over finite regions.
@@
 More generally, measures are used to define the size of sets.
-Even in probability theory, since probability measures assign probability to sets of points, in order to associate probability to individual points one needs two measures: the \textbf{probability measure} $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$.\footnote{Measures assign values to elements of a $\sigma$-algebra, which can be thought of as a set of experimental procedures.} Since we are interested in physical theories, the measure $\mu$ quantifies the number of states.
+Even in probability theory, since probability measures assign probability to sets of points, in order to associate probability to individual points one needs two measures: the \textbf{probability measure} $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$.\footnote{Measures assign values to elements of a $\sigma$-algebra, which can be thought of as a set of experimental procedures.
+\cite{aop-book}} Since we are interested in physical theories, the measure $\mu$ quantifies the number of states.
 Given a set $U$, the ratio $\frac{p(U)}{\mu(U)}$ gives us the average probability per unit of states.
@@
 However, the structure of classical phase space (i.e.
-symplectic structure) is exactly the structure that allows one to define a quantifying measure for states that is invariant under change of frame.(TODO cite book, two phil papers) The role of the quantifying measure in classical mechanics, then, is assigned to the Liouville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
+symplectic structure) is exactly the structure that allows one to define a quantifying measure for states that is invariant under change of frame.\cite{aop-book,aop-phil-HamiltonianPrivilege,aop-phil-Hamiltonianinformation} The role of the quantifying measure in classical mechanics, then, is assigned to the Liouville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
 canonical coordinates).\footnote{The volume of an infinitesimal parallelepiped is given by $dq^1 dq^2 \cdots dp_1 dp_2 \cdots$.
@@
 
-There is another expression for entropy, often referred to as the fundamental postulate of statistical mechanics, which is the logarithm of the count of microstates: $\log \mu(U)$.\footnote{Note that the Boltzmann constant is left out, as entropy will always be expressed in bits or nats.
+There is another expression for entropy, often referred to as the fundamental postulate of statistical mechanics, \cite{Peliti2011} which is the logarithm of the count of microstates: $\log \mu(U)$.\footnote{Note that the Boltzmann constant is left out, as entropy will always be expressed in bits or nats.
 Physically, the Boltzmann constant is needed to define temperature in Kelvin.
-Statistical mechanics can be equivalently formulated in terms of $\beta = \frac{1}{k_B T}$.TODO: cite the unit definition paper} In classical statistical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Liouville volume in the continuous case.
+Statistical mechanics can be equivalently formulated in terms of $\beta = \frac{1}{k_B T}$.\cite{chyla2011evolution}} In classical statistical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Liouville volume in the continuous case.
 The two expressions for entropy are linked as the Shannon/Gibbs entropy over a uniform distribution matches the logarithm of the measure of the support.
@@
 A similar connection can be established in information theory.
-The Shannon/Gibbs/von Neumann entropy quantifies the minimum number of bits required on average to send a message from a given source.\footnote{For a more general characterization of entropy that works in statistical mechanics, information theory and other contexts, and a more precise discussion of the continuous case, see TODO: cite our paper.} This, again, is a quantity that leads to empirical consequences.
+The Shannon/Gibbs/von Neumann entropy quantifies the minimum number of bits required on average to send a message from a given source.\footnote{For a more general characterization of entropy that works in statistical mechanics, information theory and other contexts, and a more precise discussion of the continuous case, see \cite{aop-phys-variability}.} This, again, is a quantity that leads to empirical consequences.
 If we have $n$ bits at our disposal, we have $2^n$ possible combinations, possible messages, that can be encoded.
@@
 So, in light of this, what is quantization?
-If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see TODO libro} If the regions are large enough, the Liouville (or Lebesgue) measure will work.
+If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see \cite{aop-book}} If the regions are large enough, the Liouville (or Lebesgue) measure will work.
 But as we shrink a region we will encounter, given the additivity, regions that contain less than one state, which does not make sense.
@@
 
-While we have found some literature on non-additive measures and set functions (CITE), together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increases.
+While we have found some literature on non-additive measures and set functions \cite{hamhalter2013quantum, nonadditive2016}, together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increases.
 Unfortunately, the quantized measure does not have this feature.
@@
 
+\section*{Acknowledgments}
+This paper is part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+This article was made possible through the support of grant \#62847 from the John Templeton Foundation.
+
 \bibliography{bibliography}

### 2023-10-28 Proceedings/2023-QIP/QuantumMeasure.tex
Fix abstract and fix/add references

@@
 	Measure theory is used in physics, not just to capture classical probability, but also to quantify the number of states.
-In previous works, in fact, we found that state quantification plays a foundational role in classical mechanics, and, therefore, we set ourselves to construct the quantum equivalent of the Liouville measure.
+In previous works, we found that state quantification plays a foundational role in classical mechanics, and therefore, we set ourselves to construct the quantum equivalent of the Liouville measure.
 Unlike the classical counterpart, this quantized measure is non-additive and has a unitary lower bound (i.e.
-no set of states cannot have less than one state).
-Conversely, requiring that state quantification is finite for finite continuous regions and unitary for each state already implies non-additivity, which implies the failure of classical theory.
-In this article we show these preliminary results and outline a new line of inquiry that may provide a different insights into the foundations of quantum theory.
-Additionally, this new approach may turn out to be useful to those interested in a quantized theory of space-time, as we believe this requires a quantized measure for the quantification of the independent degrees of freedom.
+no set of states can have less than one state).
+Conversely, requiring that state quantification is finite for finite continuous regions and that each state counts as one already implies non-additivity, which in turn implies the failure of classical theory.
+In this article we show these preliminary results and outline a new line of inquiry that may provide a different insight into the foundations of quantum theory.
+Additionally, this new approach may prove to be useful to those interested in a quantized theory of space-time, as we believe this requires a quantized measure for the quantification of the independent degrees of freedom.
 \end{abstract}
@@
 non-Kolmogorov) probability theory.
-To obviate this problem, various alternatives and extensions to measure theory and probability theory have been developed with different goals.\cite{groenewold1946principles, moyal1949quantum, gleason1975measures, sorkin1994quantum, hamhalter2013quantum, svozil2022extending, monchietti2023measure} Our interest in measure theory, however, is not in its use to capture classical probability.
+To obviate this problem, various alternatives and extensions to measure theory and probability theory have been developed with different goals.\cite{groenewold1946principles, moyal1949quantum, gleason1957measures, sorkin1994quantum, hamhalter2003quantum, gudder2009quantum, svozil2022extending, monchietti2023measure} Our interest in measure theory, however, is not in its use to capture classical probability.
 Measures play another crucial use in physical theories: ``counting'', or better ``quantifying'', states.
@@
 As part of our overall project Assumptions of Physics,\cite{aop-book} we have developed an approach, called reverse physics,\cite{aop-phys-ReversePhysics} which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive them.
-We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.\cite{aop-phil-HamiltonianPrivilege,aop-phil-Hamiltonianinformation} The former (i.e.
+We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.\cite{aop-phil-Hamiltonianinformation,aop-phil-HamiltonianPrivilege} The former (i.e.
 the symplectic structure) is the only structure that allows one to quantify states and configurations over independent degrees of freedom in a way that is frame independent.
@@
 However, this term is not widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature.
-\cite{nonadditive2016, denneberg1994non}} In section IV we will see that this non-additivity is linked to non-contextuality.
+\cite{denneberg1994non,nonadditive2014}} In section IV we will see that this non-additivity is linked to non-contextuality.
 In section V we will see that quantization means putting a unitary lower bound on the quantification of states, while keeping a finite value over finite regions.
@@
 However, the structure of classical phase space (i.e.
-symplectic structure) is exactly the structure that allows one to define a quantifying measure for states that is invariant under change of frame.\cite{aop-book,aop-phil-HamiltonianPrivilege,aop-phil-Hamiltonianinformation} The role of the quantifying measure in classical mechanics, then, is assigned to the Liouville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
+symplectic structure) is exactly the structure that allows one to define a quantifying measure for states that is invariant under change of frame.\cite{aop-book,aop-phil-Hamiltonianinformation,aop-phil-HamiltonianPrivilege} The role of the quantifying measure in classical mechanics, then, is assigned to the Liouville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
 canonical coordinates).\footnote{The volume of an infinitesimal parallelepiped is given by $dq^1 dq^2 \cdots dp_1 dp_2 \cdots$.
@@
 So, in light of this, what is quantization?
-If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see \cite{aop-book}} If the regions are large enough, the Liouville (or Lebesgue) measure will work.
+If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see \cite{kelly1996,aop-book}.} If the regions are large enough, the Liouville (or Lebesgue) measure will work.
 But as we shrink a region we will encounter, given the additivity, regions that contain less than one state, which does not make sense.
@@
 
-While we have found some literature on non-additive measures and set functions \cite{hamhalter2013quantum, nonadditive2016}, together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increases.
+While we have found some literature on non-additive measures and set functions \cite{hamhalter2003quantum, nonadditive2014}, together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increases.
 Unfortunately, the quantized measure does not have this feature.

### 2023-10-31 Proceedings/2023-QIP/Definitions/journalnames.tex
Adding mdpi template and formatted version

%%Updated on 13 October 2023
%,% Website Name,%,%,% Website Name,%%,% Full Name,%%,% Short Name,%%,% DOI abbr,%%,% ISSN,% \continuouspages
%\DeclareOption{}{\gdef\@journal{}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{acoustics}{\gdef\@journal{acoustics}\gdef\@journalfull{Acoustics}\gdef\@journalshort{Acoustics}\gdef\@doiabbr{acoustics}\gdef\@ISSN{2624-599X}\gdef\@continuouspages{yes}}
\DeclareOption{actuators}{\gdef\@journal{actuators}\gdef\@journalfull{Actuators}\gdef\@journalshort{Actuators}\gdef\@doiabbr{act}\gdef\@ISSN{2076-0825}}
\DeclareOption{addictions}{\gdef\@journal{addictions}\gdef\@journalfull{Addictions}\gdef\@journalshort{Addictions}\gdef\@doiabbr{}\gdef\@ISSN{0006-0006}}
\DeclareOption{adhesives}{\gdef\@journal{adhesives}\gdef\@journalfull{Adhesives}\gdef\@journalshort{Adhesives}\gdef\@doiabbr{adhesives}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{admsci}{\gdef\@journal{admsci}\gdef\@journalfull{Administrative Sciences}\gdef\@journalshort{Adm. Sci.}\gdef\@doiabbr{admsci}\gdef\@ISSN{2076-3387}}
\DeclareOption{adolescents}{\gdef\@journal{adolescents}\gdef\@journalfull{Adolescents}\gdef\@journalshort{Adolescents}\gdef\@doiabbr{adolescents}\gdef\@ISSN{2673-7051}\gdef\@continuouspages{yes}}
\DeclareOption{aerobiology}{\gdef\@journal{aerobiology}\gdef\@journalfull{Aerobiology}\gdef\@journalshort{Aerobiology}\gdef\@doiabbr{aerobiology}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{aerospace}{\gdef\@journal{aerospace}\gdef\@journalfull{Aerospace}\gdef\@journalshort{Aerospace}\gdef\@doiabbr{aerospace}\gdef\@ISSN{2226-4310}}
\DeclareOption{agriculture}{\gdef\@journal{agriculture}\gdef\@journalfull{Agriculture}\gdef\@journalshort{Agriculture}\gdef\@doiabbr{agriculture}\gdef\@ISSN{2077-0472}}
\DeclareOption{agriengineering}{\gdef\@journal{agriengineering}\gdef\@journalfull{AgriEngineering}\gdef\@journalshort{AgriEngineering}\gdef\@doiabbr{agriengineering}\gdef\@ISSN{2624-7402}\gdef\@continuouspages{yes}}
\DeclareOption{agrochemicals}{\gdef\@journal{agrochemicals}\gdef\@journalfull{Agrochemicals}\gdef\@journalshort{Agrochemicals}\gdef\@doiabbr{agrochemicals}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{agronomy}{\gdef\@journal{agronomy}\gdef\@journalfull{Agronomy}\gdef\@journalshort{Agronomy}\gdef\@doiabbr{agronomy}\gdef\@ISSN{2073-4395}}
\DeclareOption{ai}{\gdef\@journal{ai}\gdef\@journalfull{AI}\gdef\@journalshort{AI}\gdef\@doiabbr{ai}\gdef\@ISSN{2673-2688}\gdef\@continuouspages{yes}}
\DeclareOption{air}{\gdef\@journal{air}\gdef\@journalfull{Air}\gdef\@journalshort{Air}\gdef\@doiabbr{air}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{algorithms}{\gdef\@journal{algorithms}\gdef\@journalfull{Algorithms}\gdef\@journalshort{Algorithms}\gdef\@doiabbr{a}\gdef\@ISSN{1999-4893}}
\DeclareOption{allergies}{\gdef\@journal{allergies}\gdef\@journalfull{Allergies}\gdef\@journalshort{Allergies}\gdef\@doiabbr{allergies}\gdef\@ISSN{2313-5786}}
\DeclareOption{alloys}{\gdef\@journal{alloys}\gdef\@journalfull{Alloys}\gdef\@journalshort{Alloys}\gdef\@doiabbr{alloys}\gdef\@ISSN{2674-063X}\gdef\@continuouspages{yes}}
\DeclareOption{analytica}{\gdef\@journal{analytica}\gdef\@journalfull{Analytica}\gdef\@journalshort{Analytica}\gdef\@doiabbr{analytica}\gdef\@ISSN{2673-4532}\gdef\@continuouspages{yes}}
\DeclareOption{analytics}{\gdef\@journal{analytics}\gdef\@journalfull{Analytics}\gdef\@journalshort{Analytics}\gdef\@doiabbr{analytics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{anatomia}{\gdef\@journal{anatomia}\gdef\@journalfull{Anatomia}\gdef\@journalshort{Anatomia}\gdef\@doiabbr{anatomia}\gdef\@ISSN{2813-0545}\gdef\@continuouspages{yes}}
\DeclareOption{anesthres}{\gdef\@journal{anesthres}\gdef\@journalfull{Anesthesia Research}\gdef\@journalshort{Anesth. Res.}\gdef\@doiabbr{anesthres}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{animals}{\gdef\@journal{animals}\gdef\@journalfull{Animals}\gdef\@journalshort{Animals}\gdef\@doiabbr{ani}\gdef\@ISSN{2076-2615}}
\DeclareOption{antibiotics}{\gdef\@journal{antibiotics}\gdef\@journalfull{Antibiotics}\gdef\@journalshort{Antibiotics}\gdef\@doiabbr{antibiotics}\gdef\@ISSN{2079-6382}}
\DeclareOption{antibodies}{\gdef\@journal{antibodies}\gdef\@journalfull{Antibodies}\gdef\@journalshort{Antibodies}\gdef\@doiabbr{antib}\gdef\@ISSN{2073-4468}}
\DeclareOption{antioxidants}{\gdef\@journal{antioxidants}\gdef\@journalfull{Antioxidants}\gdef\@journalshort{Antioxidants}\gdef\@doiabbr{antiox}\gdef\@ISSN{2076-3921}}
\DeclareOption{applbiosci}{\gdef\@journal{applbiosci}\gdef\@journalfull{Applied Biosciences}\gdef\@journalshort{Appl. Biosci.}\gdef\@doiabbr{applbiosci}\gdef\@ISSN{2813-0464}\gdef\@continuouspages{yes}}
\DeclareOption{appliedchem}{\gdef\@journal{appliedchem}\gdef\@journalfull{AppliedChem}\gdef\@journalshort{AppliedChem}\gdef\@doiabbr{appliedchem}\gdef\@ISSN{2673-9623}\gdef\@continuouspages{yes}}
\DeclareOption{appliedmath}{\gdef\@journal{appliedmath}\gdef\@journalfull{AppliedMath}\gdef\@journalshort{AppliedMath}\gdef\@doiabbr{appliedmath}\gdef\@ISSN{2673-9909}\gdef\@continuouspages{yes}}
\DeclareOption{applmech}{\gdef\@journal{applmech}\gdef\@journalfull{Applied Mechanics}\gdef\@journalshort{Appl. Mech.}\gdef\@doiabbr{applmech}\gdef\@ISSN{2673-3161}\gdef\@continuouspages{yes}}
\DeclareOption{applmicrobiol}{\gdef\@journal{applmicrobiol}\gdef\@journalfull{Applied Microbiology}\gdef\@journalshort{Appl. Microbiol.}\gdef\@doiabbr{applmicrobiol}\gdef\@ISSN{2673-8007}\gdef\@continuouspages{yes}}
\DeclareOption{applnano}{\gdef\@journal{applnano}\gdef\@journalfull{Applied Nano}\gdef\@journalshort{Appl. Nano}\gdef\@doiabbr{applnano}\gdef\@ISSN{2673-3501}\gdef\@continuouspages{yes}}
\DeclareOption{applsci}{\gdef\@journal{applsci}\gdef\@journalfull{Applied Sciences}\gdef\@journalshort{Appl. Sci.}\gdef\@doiabbr{app}\gdef\@ISSN{2076-3417}}
\DeclareOption{aquacj}{\gdef\@journal{aquacj}\gdef\@journalfull{Aquaculture Journal}\gdef\@journalshort{Aquac. J.}\gdef\@doiabbr{aquacj}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{architecture}{\gdef\@journal{architecture}\gdef\@journalfull{Architecture}\gdef\@journalshort{Architecture}\gdef\@doiabbr{architecture}\gdef\@ISSN{2673-8945}\gdef\@continuouspages{yes}}
\DeclareOption{arm}{\gdef\@journal{arm}\gdef\@journalfull{Advances in Respiratory Medicine}\gdef\@journalshort{Adv. Respir. Med.}\gdef\@doiabbr{arm}\gdef\@ISSN{2543-6031}\gdef\@continuouspages{yes}}
\DeclareOption{arthropoda}{\gdef\@journal{arthropoda}\gdef\@journalfull{Arthropoda}\gdef\@journalshort{Arthropoda}\gdef\@doiabbr{arthropoda}\gdef\@ISSN{2813-3323}\gdef\@continuouspages{yes}}
\DeclareOption{arts}{\gdef\@journal{arts}\gdef\@journalfull{Arts}\gdef\@journalshort{Arts}\gdef\@doiabbr{arts}\gdef\@ISSN{2076-0752}}
\DeclareOption{asc}{\gdef\@journal{asc}\gdef\@journalfull{Automatic Systems and Control}\gdef\@journalshort{Autom. Syst. Control}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{asi}{\gdef\@journal{asi}\gdef\@journalfull{Applied System Innovation}\gdef\@journalshort{Appl. Syst. Innov.}\gdef\@doiabbr{asi}\gdef\@ISSN{2571-5577}}
\DeclareOption{astronomy}{\gdef\@journal{astronomy}\gdef\@journalfull{Astronomy}\gdef\@journalshort{Astronomy}\gdef\@doiabbr{astronomy}\gdef\@ISSN{2674-0346}\gdef\@continuouspages{yes}}
\DeclareOption{atmosphere}{\gdef\@journal{atmosphere}\gdef\@journalfull{Atmosphere}\gdef\@journalshort{Atmosphere}\gdef\@doiabbr{atmos}\gdef\@ISSN{2073-4433}}
\DeclareOption{atoms}{\gdef\@journal{atoms}\gdef\@journalfull{Atoms}\gdef\@journalshort{Atoms}\gdef\@doiabbr{atoms}\gdef\@ISSN{2218-2004}}
\DeclareOption{audiolres}{\gdef\@journal{audiolres}\gdef\@journalfull{Audiology Research}\gdef\@journalshort{Audiol. Res.}\gdef\@doiabbr{audiolres}\gdef\@ISSN{2039-4349}\gdef\@continuouspages{yes}}
\DeclareOption{automation}{\gdef\@journal{automation}\gdef\@journalfull{Automation}\gdef\@journalshort{Automation}\gdef\@doiabbr{automation}\gdef\@ISSN{2673-4052}\gdef\@continuouspages{yes}}
\DeclareOption{axioms}{\gdef\@journal{axioms}\gdef\@journalfull{Axioms}\gdef\@journalshort{Axioms}\gdef\@doiabbr{axioms}\gdef\@ISSN{2075-1680}}
\DeclareOption{bacteria}{\gdef\@journal{bacteria}\gdef\@journalfull{Bacteria}\gdef\@journalshort{Bacteria}\gdef\@doiabbr{bacteria}\gdef\@ISSN{2674-1334}\gdef\@continuouspages{yes}}
\DeclareOption{batteries}{\gdef\@journal{batteries}\gdef\@journalfull{Batteries}\gdef\@journalshort{Batteries}\gdef\@doiabbr{batteries}\gdef\@ISSN{2313-0105}}
\DeclareOption{bdcc}{\gdef\@journal{bdcc}\gdef\@journalfull{Big Data and Cognitive Computing}\gdef\@journalshort{Big Data Cogn. Comput.}\gdef\@doiabbr{bdcc}\gdef\@ISSN{2504-2289}}
\DeclareOption{behavsci}{\gdef\@journal{behavsci}\gdef\@journalfull{Behavioral Sciences}\gdef\@journalshort{Behav. Sci.}\gdef\@doiabbr{bs}\gdef\@ISSN{2076-328X}}
\DeclareOption{beverages}{\gdef\@journal{beverages}\gdef\@journalfull{Beverages}\gdef\@journalshort{Beverages}\gdef\@doiabbr{beverages}\gdef\@ISSN{2306-5710}}
\DeclareOption{biochem}{\gdef\@journal{biochem}\gdef\@journalfull{BioChem}\gdef\@journalshort{BioChem}\gdef\@doiabbr{biochem}\gdef\@ISSN{2673-6411}\gdef\@continuouspages{yes}}
\DeclareOption{bioengineering}{\gdef\@journal{bioengineering}\gdef\@journalfull{Bioengineering}\gdef\@journalshort{Bioengineering}\gdef\@doiabbr{bioengineering}\gdef\@ISSN{2306-5354}}
\DeclareOption{biologics}{\gdef\@journal{biologics}\gdef\@journalfull{Biologics}\gdef\@journalshort{Biologics}\gdef\@doiabbr{biologics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biology}{\gdef\@journal{biology}\gdef\@journalfull{Biology}\gdef\@journalshort{Biology}\gdef\@doiabbr{biology}\gdef\@ISSN{2079-7737}}
\DeclareOption{biomass}{\gdef\@journal{biomass}\gdef\@journalfull{Biomass}\gdef\@journalshort{Biomass}\gdef\@doiabbr{biomass}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biomechanics}{\gdef\@journal{biomechanics}\gdef\@journalfull{Biomechanics}\gdef\@journalshort{Biomechanics}\gdef\@doiabbr{biomechanics}\gdef\@ISSN{2673-7078}\gdef\@continuouspages{yes}}
\DeclareOption{biomed}{\gdef\@journal{biomed}\gdef\@journalfull{BioMed}\gdef\@journalshort{BioMed}\gdef\@doiabbr{biomed}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{biomedicines}{\gdef\@journal{biomedicines}\gdef\@journalfull{Biomedicines}\gdef\@journalshort{Biomedicines}\gdef\@doiabbr{biomedicines}\gdef\@ISSN{2227-9059}}
\DeclareOption{biomedinformatics}{\gdef\@journal{biomedinformatics}\gdef\@journalfull{BioMedInformatics}\gdef\@journalshort{BioMedInformatics}\gdef\@doiabbr{biomedinformatics}\gdef\@ISSN{2673-7426}\gdef\@continuouspages{yes}}
\DeclareOption{biomimetics}{\gdef\@journal{biomimetics}\gdef\@journalfull{Biomimetics}\gdef\@journalshort{Biomimetics}\gdef\@doiabbr{biomimetics}\gdef\@ISSN{2313-7673}}
\DeclareOption{biomolecules}{\gdef\@journal{biomolecules}\gdef\@journalfull{Biomolecules}\gdef\@journalshort{Biomolecules}\gdef\@doiabbr{biom}\gdef\@ISSN{2218-273X}}
\DeclareOption{biophysica}{\gdef\@journal{biophysica}\gdef\@journalfull{Biophysica}\gdef\@journalshort{Biophysica}\gdef\@doiabbr{biophysica}\gdef\@ISSN{2673-4125}\gdef\@continuouspages{yes}}
\DeclareOption{biosensors}{\gdef\@journal{biosensors}\gdef\@journalfull{Biosensors}\gdef\@journalshort{Biosensors}\gdef\@doiabbr{bios}\gdef\@ISSN{2079-6374}}
\DeclareOption{biotech}{\gdef\@journal{biotech}\gdef\@journalfull{BioTech}\gdef\@journalshort{BioTech}\gdef\@doiabbr{biotech}\gdef\@ISSN{2673-6284}}
\DeclareOption{birds}{\gdef\@journal{birds}\gdef\@journalfull{Birds}\gdef\@journalshort{Birds}\gdef\@doiabbr{birds}\gdef\@ISSN{2673-6004}\gdef\@continuouspages{yes}}
\DeclareOption{blockchains}{\gdef\@journal{blockchains}\gdef\@journalfull{Blockchains}\gdef\@journalshort{Blockchains}\gdef\@doiabbr{blockchains}\gdef\@ISSN{2813-5288}\gdef\@continuouspages{yes}}
\DeclareOption{bloods}{\gdef\@journal{bloods}\gdef\@journalfull{Bloods}\gdef\@journalshort{Bloods}\gdef\@doiabbr{bloods}\gdef\@ISSN{2673-415X}}
\DeclareOption{blsf}{\gdef\@journal{blsf}\gdef\@journalfull{Biology and Life Sciences Forum}\gdef\@journalshort{Biol. Life Sci. Forum}\gdef\@doiabbr{blsf	}\gdef\@ISSN{2673-9976}}
\DeclareOption{brainsci}{\gdef\@journal{brainsci}\gdef\@journalfull{Brain Sciences}\gdef\@journalshort{Brain Sci.}\gdef\@doiabbr{brainsci}\gdef\@ISSN{2076-3425}}
\DeclareOption{breath}{\gdef\@journal{breath}\gdef\@journalfull{Breath}\gdef\@journalshort{Breath}\gdef\@doiabbr{breath}\gdef\@ISSN{}}
\DeclareOption{buildings}{\gdef\@journal{buildings}\gdef\@journalfull{Buildings}\gdef\@journalshort{Buildings}\gdef\@doiabbr{buildings}\gdef\@ISSN{2075-5309}}
\DeclareOption{businesses}{\gdef\@journal{businesses}\gdef\@journalfull{Businesses}\gdef\@journalshort{Businesses}\gdef\@doiabbr{businesses}\gdef\@ISSN{2673-7116}\gdef\@continuouspages{yes}}
\DeclareOption{cancers}{\gdef\@journal{cancers}\gdef\@journalfull{Cancers}\gdef\@journalshort{Cancers}\gdef\@doiabbr{cancers}\gdef\@ISSN{2072-6694}}
\DeclareOption{carbon}{\gdef\@journal{carbon}\gdef\@journalfull{C}\gdef\@journalshort{C}\gdef\@doiabbr{c}\gdef\@ISSN{2311-5629}}
\DeclareOption{cardiogenetics}{\gdef\@journal{cardiogenetics}\gdef\@journalfull{Cardiogenetics}\gdef\@journalshort{Cardiogenetics}\gdef\@doiabbr{cardiogenetics}\gdef\@ISSN{2035-8148}\gdef\@continuouspages{yes}}
\DeclareOption{catalysts}{\gdef\@journal{catalysts}\gdef\@journalfull{Catalysts}\gdef\@journalshort{Catalysts}\gdef\@doiabbr{catal}\gdef\@ISSN{2073-4344}}
\DeclareOption{cells}{\gdef\@journal{cells}\gdef\@journalfull{Cells}\gdef\@journalshort{Cells}\gdef\@doiabbr{cells}\gdef\@ISSN{2073-4409}}
\DeclareOption{ceramics}{\gdef\@journal{ceramics}\gdef\@journalfull{Ceramics}\gdef\@journalshort{Ceramics}\gdef\@doiabbr{ceramics}\gdef\@ISSN{2571-6131}\gdef\@continuouspages{yes}}
\DeclareOption{challenges}{\gdef\@journal{challenges}\gdef\@journalfull{Challenges}\gdef\@journalshort{Challenges}\gdef\@doiabbr{challe}\gdef\@ISSN{2078-1547}}
\DeclareOption{chemengineering}{\gdef\@journal{chemengineering}\gdef\@journalfull{ChemEngineering}\gdef\@journalshort{ChemEngineering}\gdef\@doiabbr{chemengineering}\gdef\@ISSN{2305-7084}}
\DeclareOption{chemistry}{\gdef\@journal{chemistry}\gdef\@journalfull{Chemistry}\gdef\@journalshort{Chemistry}\gdef\@doiabbr{chemistry}\gdef\@ISSN{2624-8549}\gdef\@continuouspages{yes}}
\DeclareOption{chemosensors}{\gdef\@journal{chemosensors}\gdef\@journalfull{Chemosensors}\gdef\@journalshort{Chemosensors}\gdef\@doiabbr{chemosensors}\gdef\@ISSN{2227-9040}}
\DeclareOption{chemproc}{\gdef\@journal{chemproc}\gdef\@journalfull{Chemistry Proceedings}\gdef\@journalshort{Chem. Proc.}\gdef\@doiabbr{chemproc}\gdef\@ISSN{2673-4583}}
\DeclareOption{children}{\gdef\@journal{children}\gdef\@journalfull{Children}\gdef\@journalshort{Children}\gdef\@doiabbr{children}\gdef\@ISSN{2227-9067}}
\DeclareOption{chips}{\gdef\@journal{chips}\gdef\@journalfull{Chips}\gdef\@journalshort{Chips}\gdef\@doiabbr{chips}\gdef\@ISSN{2674-0710}\gdef\@continuouspages{yes}}
\DeclareOption{cimb}{\gdef\@journal{cimb}\gdef\@journalfull{Current Issues in Molecular Biology}\gdef\@journalshort{Curr. Issues Mol. Biol.}\gdef\@doiabbr{cimb}\gdef\@ISSN{1467-3045}\gdef\@continuouspages{yes}}
\DeclareOption{civileng}{\gdef\@journal{civileng}\gdef\@journalfull{CivilEng}\gdef\@journalshort{CivilEng}\gdef\@doiabbr{civileng}\gdef\@ISSN{2673-4109}\gdef\@continuouspages{yes}}
\DeclareOption{cleantechnol}{\gdef\@journal{cleantechnol}\gdef\@journalfull{Clean Technologies}\gdef\@journalshort{Clean Technol.}\gdef\@doiabbr{cleantechnol}\gdef\@ISSN{2571-8797}\gdef\@continuouspages{yes}}
\DeclareOption{climate}{\gdef\@journal{climate}\gdef\@journalfull{Climate}\gdef\@journalshort{Climate}\gdef\@doiabbr{cli}\gdef\@ISSN{2225-1154}}
\DeclareOption{clinpract}{\gdef\@journal{clinpract}\gdef\@journalfull{Clinics and Practice}\gdef\@journalshort{Clin. Pract.}\gdef\@doiabbr{clinpract}\gdef\@ISSN{2039-7283}\gdef\@continuouspages{yes}}
\DeclareOption{clockssleep}{\gdef\@journal{clockssleep}\gdef\@journalfull{Clocks \& Sleep}\gdef\@journalshort{Clocks\&Sleep}\gdef\@doiabbr{clockssleep}\gdef\@ISSN{2624-5175}\gdef\@continuouspages{yes}}
\DeclareOption{cmd}{\gdef\@journal{cmd}\gdef\@journalfull{Corrosion and Materials Degradation}\gdef\@journalshort{Corros. Mater. Degrad.}\gdef\@doiabbr{cmd}\gdef\@ISSN{2624-5558}\gdef\@continuouspages{yes}}
\DeclareOption{coasts}{\gdef\@journal{coasts}\gdef\@journalfull{Coasts}\gdef\@journalshort{Coasts}\gdef\@doiabbr{coasts}\gdef\@ISSN{2673-964X}\gdef\@continuouspages{yes}}
\DeclareOption{coatings}{\gdef\@journal{coatings}\gdef\@journalfull{Coatings}\gdef\@journalshort{Coatings}\gdef\@doiabbr{coatings}\gdef\@ISSN{2079-6412}}
\DeclareOption{colloids}{\gdef\@journal{colloids}\gdef\@journalfull{Colloids and Interfaces}\gdef\@journalshort{Colloids Interfaces}\gdef\@doiabbr{colloids}\gdef\@ISSN{2504-5377}}
\DeclareOption{colorants}{\gdef\@journal{colorants}\gdef\@journalfull{Colorants}\gdef\@journalshort{Colorants}\gdef\@doiabbr{colorants}\gdef\@ISSN{2079-6447}\gdef\@continuouspages{yes}}
\DeclareOption{commodities}{\gdef\@journal{commodities}\gdef\@journalfull{Commodities}\gdef\@journalshort{Commodities}\gdef\@doiabbr{commodities}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{compounds}{\gdef\@journal{compounds}\gdef\@journalfull{Compounds}\gdef\@journalshort{Compounds}\gdef\@doiabbr{compounds}\gdef\@ISSN{2673-6918}\gdef\@continuouspages{yes}}
\DeclareOption{computation}{\gdef\@journal{computation}\gdef\@journalfull{Computation}\gdef\@journalshort{Computation}\gdef\@doiabbr{computation}\gdef\@ISSN{2079-3197}}
\DeclareOption{computers}{\gdef\@journal{computers}\gdef\@journalfull{Computers}\gdef\@journalshort{Computers}\gdef\@doiabbr{computers}\gdef\@ISSN{2073-431X}}
\DeclareOption{condensedmatter}{\gdef\@journal{condensedmatter}\gdef\@journalfull{Condensed Matter}\gdef\@journalshort{Condens. Matter}\gdef\@doiabbr{condmat}\gdef\@ISSN{2410-3896}}
\DeclareOption{conservation}{\gdef\@journal{conservation}\gdef\@journalfull{Conservation}\gdef\@journalshort{Conservation}\gdef\@doiabbr{conservation}\gdef\@ISSN{2673-7159}\gdef\@continuouspages{yes}}
\DeclareOption{constrmater}{\gdef\@journal{constrmater}\gdef\@journalfull{Construction Materials}\gdef\@journalshort{Constr. Mater.}\gdef\@doiabbr{constrmater}\gdef\@ISSN{2673-7108}\gdef\@continuouspages{yes}}
\DeclareOption{cosmetics}{\gdef\@journal{cosmetics}\gdef\@journalfull{Cosmetics}\gdef\@journalshort{Cosmetics}\gdef\@doiabbr{cosmetics}\gdef\@ISSN{2079-9284}}
\DeclareOption{covid}{\gdef\@journal{covid}\gdef\@journalfull{COVID}\gdef\@journalshort{COVID}\gdef\@doiabbr{covid}\gdef\@ISSN{2673-8112}\gdef\@continuouspages{yes}}
\DeclareOption{crops}{\gdef\@journal{crops}\gdef\@journalfull{Crops}\gdef\@journalshort{Crops}\gdef\@doiabbr{crops}\gdef\@ISSN{2673-7655}\gdef\@continuouspages{yes}}
\DeclareOption{cryptography}{\gdef\@journal{cryptography}\gdef\@journalfull{Cryptography}\gdef\@journalshort{Cryptography}\gdef\@doiabbr{cryptography}\gdef\@ISSN{2410-387X}}
\DeclareOption{crystals}{\gdef\@journal{crystals}\gdef\@journalfull{Crystals}\gdef\@journalshort{Crystals}\gdef\@doiabbr{cryst}\gdef\@ISSN{2073-4352}}
\DeclareOption{csmf}{\gdef\@journal{csmf}\gdef\@journalfull{Computer Sciences \& Mathematics Forum}\gdef\@journalshort{Comput. Sci. Math. Forum}\gdef\@doiabbr{csmf}\gdef\@ISSN{2813-0324}}
\DeclareOption{ctn}{\gdef\@journal{ctn}\gdef\@journalfull{Clinical and Translational Neuroscience}\gdef\@journalshort{Clin. Transl. Neurosci.}\gdef\@doiabbr{ctn}\gdef\@ISSN{2514-183X}}
\DeclareOption{curroncol}{\gdef\@journal{curroncol}\gdef\@journalfull{Current Oncology}\gdef\@journalshort{Curr. Oncol.}\gdef\@doiabbr{curroncol}\gdef\@ISSN{1718-7729}\gdef\@continuouspages{yes}}
\DeclareOption{cyber}{\gdef\@journal{cyber}\gdef\@journalfull{Cyber}\gdef\@journalshort{Cyber}\gdef\@doiabbr{cyber}\gdef\@ISSN{2624-800X}}
\DeclareOption{dairy}{\gdef\@journal{dairy}\gdef\@journalfull{Dairy}\gdef\@journalshort{Dairy}\gdef\@doiabbr{dairy}\gdef\@ISSN{2624-862X}\gdef\@continuouspages{yes}}
\DeclareOption{data}{\gdef\@journal{data}\gdef\@journalfull{Data}\gdef\@journalshort{Data}\gdef\@doiabbr{data}\gdef\@ISSN{2306-5729}}
\DeclareOption{ddc}{\gdef\@journal{ddc}\gdef\@journalfull{Drugs and Drug Candidates}\gdef\@journalshort{Drugs Drug Candidates}\gdef\@doiabbr{ddc}\gdef\@ISSN{2813-2998}\gdef\@continuouspages{yes}}
\DeclareOption{dentistry}{\gdef\@journal{dentistry}\gdef\@journalfull{Dentistry Journal}\gdef\@journalshort{Dent. J.}\gdef\@doiabbr{dj}\gdef\@ISSN{2304-6767}}
\DeclareOption{dermato}{\gdef\@journal{dermato}\gdef\@journalfull{Dermato}\gdef\@journalshort{Dermato}\gdef\@doiabbr{dermato}\gdef\@ISSN{2673-6179}\gdef\@continuouspages{yes}}
\DeclareOption{dermatopathology}{\gdef\@journal{dermatopathology}\gdef\@journalfull{Dermatopathology}\gdef\@journalshort{Dermatopathology}\gdef\@doiabbr{dermatopathology}\gdef\@ISSN{2296-3529}\gdef\@continuouspages{yes}}
\DeclareOption{designs}{\gdef\@journal{designs}\gdef\@journalfull{Designs}\gdef\@journalshort{Designs}\gdef\@doiabbr{designs}\gdef\@ISSN{2411-9660}}
\DeclareOption{devices}{\gdef\@journal{devices}\gdef\@journalfull{Devices}\gdef\@journalshort{Devices}\gdef\@doiabbr{devices}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{diabetology}{\gdef\@journal{diabetology}\gdef\@journalfull{Diabetology}\gdef\@journalshort{Diabetology}\gdef\@doiabbr{diabetology}\gdef\@ISSN{2673-4540}\gdef\@continuouspages{yes}}
\DeclareOption{diagnostics}{\gdef\@journal{diagnostics}\gdef\@journalfull{Diagnostics}\gdef\@journalshort{Diagnostics}\gdef\@doiabbr{diagnostics}\gdef\@ISSN{2075-4418}}
\DeclareOption{dietetics}{\gdef\@journal{dietetics}\gdef\@journalfull{Dietetics}\gdef\@journalshort{Dietetics}\gdef\@doiabbr{dietetics}\gdef\@ISSN{2674-0311}\gdef\@continuouspages{yes}}
\DeclareOption{digital}{\gdef\@journal{digital}\gdef\@journalfull{Digital}\gdef\@journalshort{Digital}\gdef\@doiabbr{digital}\gdef\@ISSN{2673-6470}\gdef\@continuouspages{yes}}
\DeclareOption{disabilities}{\gdef\@journal{disabilities}\gdef\@journalfull{Disabilities}\gdef\@journalshort{Disabilities}\gdef\@doiabbr{disabilities}\gdef\@ISSN{2673-7272}\gdef\@continuouspages{yes}}
\DeclareOption{diseases}{\gdef\@journal{diseases}\gdef\@journalfull{Diseases}\gdef\@journalshort{Diseases}\gdef\@doiabbr{diseases}\gdef\@ISSN{2079-9721}}
\DeclareOption{diversity}{\gdef\@journal{diversity}\gdef\@journalfull{Diversity}\gdef\@journalshort{Diversity}\gdef\@doiabbr{d}\gdef\@ISSN{1424-2818}}
\DeclareOption{dna}{\gdef\@journal{dna}\gdef\@journalfull{DNA}\gdef\@journalshort{DNA}\gdef\@doiabbr{dna}\gdef\@ISSN{2673-8856}\gdef\@continuouspages{yes}}
\DeclareOption{drones}{\gdef\@journal{drones}\gdef\@journalfull{Drones}\gdef\@journalshort{Drones}\gdef\@doiabbr{drones}\gdef\@ISSN{2504-446X}}
\DeclareOption{dynamics}{\gdef\@journal{dynamics}\gdef\@journalfull{Dynamics}\gdef\@journalshort{Dynamics}\gdef\@doiabbr{dynamics}\gdef\@ISSN{2673-8716}\gdef\@continuouspages{yes}}
\DeclareOption{earth}{\gdef\@journal{earth}\gdef\@journalfull{Earth}\gdef\@journalshort{Earth}\gdef\@doiabbr{earth}\gdef\@ISSN{2673-4834}\gdef\@continuouspages{yes}}
\DeclareOption{ebj}{\gdef\@journal{ebj}\gdef\@journalfull{European Burn Journal}\gdef\@journalshort{Eur. Burn J.}\gdef\@doiabbr{ebj}\gdef\@ISSN{2673-1991}\gdef\@continuouspages{yes}}
\DeclareOption{ecm}{\gdef\@journal{ecm}\gdef\@journalfull{Emergency Care and Medicine}\gdef\@journalshort{Emerg. Care Med.}\gdef\@doiabbr{ecm}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{ecologies}{\gdef\@journal{ecologies}\gdef\@journalfull{Ecologies}\gdef\@journalshort{Ecologies}\gdef\@doiabbr{ecologies}\gdef\@ISSN{2673-4133}\gdef\@continuouspages{yes}}
\DeclareOption{econometrics}{\gdef\@journal{econometrics}\gdef\@journalfull{Econometrics}\gdef\@journalshort{Econometrics}\gdef\@doiabbr{econometrics}\gdef\@ISSN{2225-1146}}
\DeclareOption{economies}{\gdef\@journal{economies}\gdef\@journalfull{Economies}\gdef\@journalshort{Economies}\gdef\@doiabbr{economies}\gdef\@ISSN{2227-7099}}
\DeclareOption{education}{\gdef\@journal{education}\gdef\@journalfull{Education Sciences}\gdef\@journalshort{Educ. Sci.}\gdef\@doiabbr{educsci}\gdef\@ISSN{2227-7102}}
\DeclareOption{ejihpe}{\gdef\@journal{ejihpe}\gdef\@journalfull{European Journal of Investigation in Health, Psychology and Education}\gdef\@journalshort{Eur. J. Investig. Health Psychol. Educ.}\gdef\@doiabbr{ejihpe}\gdef\@ISSN{2254-9625}\gdef\@continuouspages{yes}}
\DeclareOption{electricity}{\gdef\@journal{electricity}\gdef\@journalfull{Electricity}\gdef\@journalshort{Electricity}\gdef\@doiabbr{electricity}\gdef\@ISSN{2673-4826}\gdef\@continuouspages{yes}}
\DeclareOption{electrochem}{\gdef\@journal{electrochem}\gdef\@journalfull{Electrochem}\gdef\@journalshort{Electrochem}\gdef\@doiabbr{electrochem}\gdef\@ISSN{2673-3293}\gdef\@continuouspages{yes}}
\DeclareOption{electronicmat}{\gdef\@journal{electronicmat}\gdef\@journalfull{Electronic Materials}\gdef\@journalshort{Electron. Mater.}\gdef\@doiabbr{electronicmat}\gdef\@ISSN{2673-3978}\gdef\@continuouspages{yes}}
\DeclareOption{electronics}{\gdef\@journal{electronics}\gdef\@journalfull{Electronics}\gdef\@journalshort{Electronics}\gdef\@doiabbr{electronics}\gdef\@ISSN{2079-9292}}
\DeclareOption{encyclopedia}{\gdef\@journal{encyclopedia}\gdef\@journalfull{Encyclopedia}\gdef\@journalshort{Encyclopedia}\gdef\@doiabbr{encyclopedia}\gdef\@ISSN{2673-8392}\gdef\@continuouspages{yes}}
\DeclareOption{endocrines}{\gdef\@journal{endocrines}\gdef\@journalfull{Endocrines}\gdef\@journalshort{Endocrines}\gdef\@doiabbr{endocrines}\gdef\@ISSN{2673-396X}\gdef\@continuouspages{yes}}
\DeclareOption{energies}{\gdef\@journal{energies}\gdef\@journalfull{Energies}\gdef\@journalshort{Energies}\gdef\@doiabbr{en}\gdef\@ISSN{1996-1073}}
\DeclareOption{eng}{\gdef\@journal{eng}\gdef\@journalfull{Eng}\gdef\@journalshort{Eng}\gdef\@doiabbr{eng}\gdef\@ISSN{2673-4117}\gdef\@continuouspages{yes}}
\DeclareOption{engproc}{\gdef\@journal{engproc}\gdef\@journalfull{Engineering Proceedings}\gdef\@journalshort{Eng. Proc.}\gdef\@doiabbr{engproc}\gdef\@ISSN{2673-4591}}
%\DeclareOption{ent}{\gdef\@journal{ent}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{entomology}{\gdef\@journal{entomology}\gdef\@journalfull{Entomology}\gdef\@journalshort{Entomology}\gdef\@doiabbr{entomology}\gdef\@ISSN{2813-0502}\gdef\@continuouspages{yes}}	
\DeclareOption{entropy}{\gdef\@journal{entropy}\gdef\@journalfull{Entropy}\gdef\@journalshort{Entropy}\gdef\@doiabbr{e}\gdef\@ISSN{1099-4300}}
\DeclareOption{environments}{\gdef\@journal{environments}\gdef\@journalfull{Environments}\gdef\@journalshort{Environments}\gdef\@doiabbr{environments}\gdef\@ISSN{2076-3298}}
\DeclareOption{environsciproc}{\gdef\@journal{environsciproc}\gdef\@journalfull{Environmental Sciences Proceedings}\gdef\@journalshort{Environ. Sci. Proc.}\gdef\@doiabbr{environsciproc}\gdef\@ISSN{2673-4931}}
\DeclareOption{epidemiologia}{\gdef\@journal{epidemiologia}\gdef\@journalfull{Epidemiologia}\gdef\@journalshort{Epidemiologia}\gdef\@doiabbr{epidemiologia}\gdef\@ISSN{2673-3986}\gdef\@continuouspages{yes}}
\DeclareOption{epigenomes}{\gdef\@journal{epigenomes}\gdef\@journalfull{Epigenomes}\gdef\@journalshort{Epigenomes}\gdef\@doiabbr{epigenomes}\gdef\@ISSN{2075-4655}}
\DeclareOption{est}{\gdef\@journal{est}\gdef\@journalfull{Electrochemical Science and Technology}\gdef\@journalshort{Electrochem. Sci. Technol.}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{fermentation}{\gdef\@journal{fermentation}\gdef\@journalfull{Fermentation}\gdef\@journalshort{Fermentation}\gdef\@doiabbr{fermentation}\gdef\@ISSN{2311-5637}}
\DeclareOption{fibers}{\gdef\@journal{fibers}\gdef\@journalfull{Fibers}\gdef\@journalshort{Fibers}\gdef\@doiabbr{fib}\gdef\@ISSN{2079-6439}}
\DeclareOption{fintech}{\gdef\@journal{fintech}\gdef\@journalfull{FinTech}\gdef\@journalshort{FinTech}\gdef\@doiabbr{fintech}\gdef\@ISSN{2674-1032}\gdef\@continuouspages{yes}}
\DeclareOption{fire}{\gdef\@journal{fire}\gdef\@journalfull{Fire}\gdef\@journalshort{Fire}\gdef\@doiabbr{fire}\gdef\@ISSN{2571-6255}}
\DeclareOption{fishes}{\gdef\@journal{fishes}\gdef\@journalfull{Fishes}\gdef\@journalshort{Fishes}\gdef\@doiabbr{fishes}\gdef\@ISSN{2410-3888}}
\DeclareOption{fluids}{\gdef\@journal{fluids}\gdef\@journalfull{Fluids}\gdef\@journalshort{Fluids}\gdef\@doiabbr{fluids}\gdef\@ISSN{2311-5521}}
\DeclareOption{foods}{\gdef\@journal{foods}\gdef\@journalfull{Foods}\gdef\@journalshort{Foods}\gdef\@doiabbr{foods}\gdef\@ISSN{2304-8158}}
\DeclareOption{forecasting}{\gdef\@journal{forecasting}\gdef\@journalfull{Forecasting}\gdef\@journalshort{Forecasting}\gdef\@doiabbr{forecast}\gdef\@ISSN{2571-9394}\gdef\@continuouspages{yes}}
\DeclareOption{forensicsci}{\gdef\@journal{forensicsci}\gdef\@journalfull{Forensic Sciences}\gdef\@journalshort{Forensic Sci.}\gdef\@doiabbr{forensicsci}\gdef\@ISSN{2673-6756}\gdef\@continuouspages{yes}}
\DeclareOption{forests}{\gdef\@journal{forests}\gdef\@journalfull{Forests}\gdef\@journalshort{Forests}\gdef\@doiabbr{f}\gdef\@ISSN{1999-4907}}
\DeclareOption{fossils}{\gdef\@journal{fossils}\gdef\@journalfull{Fossils}\gdef\@journalshort{Fossils}\gdef\@doiabbr{fossils}\gdef\@ISSN{2813-6284}\gdef\@continuouspages{yes}}
\DeclareOption{foundations}{\gdef\@journal{foundations}\gdef\@journalfull{Foundations}\gdef\@journalshort{Foundations}\gdef\@doiabbr{foundations}\gdef\@ISSN{2673-9321}\gdef\@continuouspages{yes}}
\DeclareOption{fractalfract}{\gdef\@journal{fractalfract}\gdef\@journalfull{Fractal and Fractional}\gdef\@journalshort{Fractal Fract.}\gdef\@doiabbr{fractalfract}\gdef\@ISSN{2504-3110}}
\DeclareOption{fuels}{\gdef\@journal{fuels}\gdef\@journalfull{Fuels}\gdef\@journalshort{Fuels}\gdef\@doiabbr{fuels}\gdef\@ISSN{2673-3994}\gdef\@continuouspages{yes}}
\DeclareOption{future}{\gdef\@journal{future}\gdef\@journalfull{Future}\gdef\@journalshort{Future}\gdef\@doiabbr{future}\gdef\@ISSN{2813-2882}\gdef\@continuouspages{yes}}
\DeclareOption{futureinternet}{\gdef\@journal{futureinternet}\gdef\@journalfull{Future Internet}\gdef\@journalshort{Future Internet}\gdef\@doiabbr{fi}\gdef\@ISSN{1999-5903}}
%\DeclareOption{futureparasites}{\gdef\@journal{futureparasites}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{futurepharmacol}{\gdef\@journal{futurepharmacol}\gdef\@journalfull{Future Pharmacology}\gdef\@journalshort{Future Pharmacol.}\gdef\@doiabbr{futurepharmacol}\gdef\@ISSN{2673-9879}\gdef\@continuouspages{yes}}
\DeclareOption{futurephys}{\gdef\@journal{futurephys}\gdef\@journalfull{Future Physics}\gdef\@journalshort{Future Phys.}\gdef\@doiabbr{futurephys}\gdef\@ISSN{2624-6503}}
\DeclareOption{futuretransp}{\gdef\@journal{futuretransp}\gdef\@journalfull{Future Transportation}\gdef\@journalshort{Future Transp.}\gdef\@doiabbr{futuretransp}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{galaxies}{\gdef\@journal{galaxies}\gdef\@journalfull{Galaxies}\gdef\@journalshort{Galaxies}\gdef\@doiabbr{galaxies}\gdef\@ISSN{2075-4434}}
\DeclareOption{games}{\gdef\@journal{games}\gdef\@journalfull{Games}\gdef\@journalshort{Games}\gdef\@doiabbr{g}\gdef\@ISSN{2073-4336}}
\DeclareOption{gases}{\gdef\@journal{gases}\gdef\@journalfull{Gases}\gdef\@journalshort{Gases}\gdef\@doiabbr{gases}\gdef\@ISSN{2673-5628}\gdef\@continuouspages{yes}}
\DeclareOption{gastroent}{\gdef\@journal{gastroent}\gdef\@journalfull{Gastroenterology Insights}\gdef\@journalshort{Gastroenterol. Insights}\gdef\@doiabbr{gastroent}\gdef\@ISSN{2036-7422}\gdef\@continuouspages{yes}}
\DeclareOption{gastrointestdisord}{\gdef\@journal{gastrointestdisord}\gdef\@journalfull{Gastrointestinal Disorders}\gdef\@journalshort{Gastrointest. Disord.}\gdef\@doiabbr{gidisord}\gdef\@ISSN{2624-5647}\gdef\@continuouspages{yes}}
\DeclareOption{gastronomy}{\gdef\@journal{gastronomy}\gdef\@journalfull{Gastronomy}\gdef\@journalshort{Gastronomy}\gdef\@doiabbr{gastronomy}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{gels}{\gdef\@journal{gels}\gdef\@journalfull{Gels}\gdef\@journalshort{Gels}\gdef\@doiabbr{gels}\gdef\@ISSN{2310-2861}}
\DeclareOption{genealogy}{\gdef\@journal{genealogy}\gdef\@journalfull{Genealogy}\gdef\@journalshort{Genealogy}\gdef\@doiabbr{genealogy}\gdef\@ISSN{2313-5778}}
\DeclareOption{genes}{\gdef\@journal{genes}\gdef\@journalfull{Genes}\gdef\@journalshort{Genes}\gdef\@doiabbr{genes}\gdef\@ISSN{2073-4425}}
\DeclareOption{geographies}{\gdef\@journal{geographies}\gdef\@journalfull{Geographies}\gdef\@journalshort{Geographies}\gdef\@doiabbr{geographies}\gdef\@ISSN{2673-7086}\gdef\@continuouspages{yes}}
\DeclareOption{geohazards}{\gdef\@journal{geohazards}\gdef\@journalfull{GeoHazards}\gdef\@journalshort{GeoHazards}\gdef\@doiabbr{geohazards}\gdef\@ISSN{2624-795X}\gdef\@continuouspages{yes}}
\DeclareOption{geomatics}{\gdef\@journal{geomatics}\gdef\@journalfull{Geomatics}\gdef\@journalshort{Geomatics}\gdef\@doiabbr{geomatics}\gdef\@ISSN{2673-7418}\gdef\@continuouspages{yes}}
\DeclareOption{geosciences}{\gdef\@journal{geosciences}\gdef\@journalfull{Geosciences}\gdef\@journalshort{Geosciences}\gdef\@doiabbr{geosciences}\gdef\@ISSN{2076-3263}}
\DeclareOption{geotechnics}{\gdef\@journal{geotechnics}\gdef\@journalfull{Geotechnics}\gdef\@journalshort{Geotechnics}\gdef\@doiabbr{geotechnics}\gdef\@ISSN{2673-7094}\gdef\@continuouspages{yes}}
\DeclareOption{geriatrics}{\gdef\@journal{geriatrics}\gdef\@journalfull{Geriatrics}\gdef\@journalshort{Geriatrics}\gdef\@doiabbr{geriatrics}\gdef\@ISSN{2308-3417}}
\DeclareOption{grasses}{\gdef\@journal{grasses}\gdef\@journalfull{Grasses}\gdef\@journalshort{Grasses}\gdef\@doiabbr{grasses}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{gucdd}{\gdef\@journal{gucdd}\gdef\@journalfull{Gout, Urate, and Crystal Deposition Disease}\gdef\@journalshort{Gout Urate Cryst. Depos. Dis.}\gdef\@doiabbr{gucdd}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{hardware}{\gdef\@journal{hardware}\gdef\@journalfull{Hardware}\gdef\@journalshort{Hardware}\gdef\@doiabbr{hardware}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{hazardousmatters}{\gdef\@journal{hazardousmatters}\gdef\@journalfull{Hazardous Matters}\gdef\@journalshort{Hazard. Matters}\gdef\@doiabbr{}\gdef\@ISSN{0014-0014}}
\DeclareOption{healthcare}{\gdef\@journal{healthcare}\gdef\@journalfull{Healthcare}\gdef\@journalshort{Healthcare}\gdef\@doiabbr{healthcare}\gdef\@ISSN{2227-9032}}
\DeclareOption{hearts}{\gdef\@journal{hearts}\gdef\@journalfull{Hearts}\gdef\@journalshort{Hearts}\gdef\@doiabbr{hearts}\gdef\@ISSN{2673-3846}\gdef\@continuouspages{yes}}
\DeclareOption{hemato}{\gdef\@journal{hemato}\gdef\@journalfull{Hemato}\gdef\@journalshort{Hemato}\gdef\@doiabbr{hemato}\gdef\@ISSN{2673-6357}\gdef\@continuouspages{yes}}
\DeclareOption{hematolrep}{\gdef\@journal{hematolrep}\gdef\@journalfull{Hematology Reports}\gdef\@journalshort{Hematol. Rep.}\gdef\@doiabbr{hematolrep}\gdef\@ISSN{2038-8330}\gdef\@continuouspages{yes}}
\DeclareOption{heritage}{\gdef\@journal{heritage}\gdef\@journalfull{Heritage}\gdef\@journalshort{Heritage}\gdef\@doiabbr{heritage}\gdef\@ISSN{2571-9408}\gdef\@continuouspages{yes}}
\DeclareOption{higheredu}{\gdef\@journal{higheredu}\gdef\@journalfull{Trends in Higher Education}\gdef\@journalshort{Trends High. Educ.}\gdef\@doiabbr{higheredu}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{highthroughput}{\gdef\@journal{highthroughput}\gdef\@journalfull{High-Throughput}\gdef\@journalshort{High-Throughput}\gdef\@doiabbr{ht}\gdef\@ISSN{2571-5135}}
\DeclareOption{histories}{\gdef\@journal{histories}\gdef\@journalfull{Histories}\gdef\@journalshort{Histories}\gdef\@doiabbr{histories}\gdef\@ISSN{2409-9252}\gdef\@continuouspages{yes}}
\DeclareOption{horticulturae}{\gdef\@journal{horticulturae}\gdef\@journalfull{Horticulturae}\gdef\@journalshort{Horticulturae}\gdef\@doiabbr{horticulturae}\gdef\@ISSN{2311-7524}}
\DeclareOption{hospitals}{\gdef\@journal{hospitals}\gdef\@journalfull{Hospitals}\gdef\@journalshort{Hospitals}\gdef\@doiabbr{hospitals}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{humanities}{\gdef\@journal{humanities}\gdef\@journalfull{Humanities}\gdef\@journalshort{Humanities}\gdef\@doiabbr{h}\gdef\@ISSN{2076-0787}}
\DeclareOption{humans}{\gdef\@journal{humans}\gdef\@journalfull{Humans}\gdef\@journalshort{Humans}\gdef\@doiabbr{humans}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{hydrobiology}{\gdef\@journal{hydrobiology}\gdef\@journalfull{Hydrobiology}\gdef\@journalshort{Hydrobiology	}\gdef\@doiabbr{hydrobiology}\gdef\@ISSN{2673-9917}\gdef\@continuouspages{yes}}
\DeclareOption{hydrogen}{\gdef\@journal{hydrogen}\gdef\@journalfull{Hydrogen}\gdef\@journalshort{Hydrogen}\gdef\@doiabbr{hydrogen}\gdef\@ISSN{2673-4141}\gdef\@continuouspages{yes}}
\DeclareOption{hydrology}{\gdef\@journal{hydrology}\gdef\@journalfull{Hydrology}\gdef\@journalshort{Hydrology}\gdef\@doiabbr{hydrology}\gdef\@ISSN{2306-5338}}
\DeclareOption{hygiene}{\gdef\@journal{hygiene}\gdef\@journalfull{Hygiene}\gdef\@journalshort{Hygiene}\gdef\@doiabbr{hygiene}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{ice}{\gdef\@journal{ice}\gdef\@journalfull{Ice}\gdef\@journalshort{Ice}\gdef\@doiabbr{ice}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{idr}{\gdef\@journal{idr}\gdef\@journalfull{Infectious Disease Reports}\gdef\@journalshort{Infect. Dis. Rep.}\gdef\@doiabbr{idr}\gdef\@ISSN{2036-7449}\gdef\@continuouspages{yes}}
\DeclareOption{ijerph}{\gdef\@journal{ijerph}\gdef\@journalfull{International Journal of Environmental Research and Public Health}\gdef\@journalshort{Int. J. Environ. Res. Public Health}\gdef\@doiabbr{ijerph}\gdef\@ISSN{1660-4601}}
\DeclareOption{ijfs}{\gdef\@journal{ijfs}\gdef\@journalfull{International Journal of Financial Studies}\gdef\@journalshort{Int. J. Financial Stud.}\gdef\@doiabbr{ijfs}\gdef\@ISSN{2227-7072}}
\DeclareOption{ijgi}{\gdef\@journal{ijgi}\gdef\@journalfull{ISPRS International Journal of Geo-Information}\gdef\@journalshort{ISPRS Int. J. Geo-Inf.}\gdef\@doiabbr{ijgi}\gdef\@ISSN{2220-9964}}
\DeclareOption{ijms}{\gdef\@journal{ijms}\gdef\@journalfull{International Journal of Molecular Sciences}\gdef\@journalshort{Int. J. Mol. Sci.}\gdef\@doiabbr{ijms}\gdef\@ISSN{1422-0067}}
\DeclareOption{ijns}{\gdef\@journal{ijns}\gdef\@journalfull{International Journal of Neonatal Screening}\gdef\@journalshort{Int. J. Neonatal Screen.}\gdef\@doiabbr{ijns}\gdef\@ISSN{2409-515X}}
\DeclareOption{ijpb}{\gdef\@journal{ijpb}\gdef\@journalfull{International Journal of Plant Biology}\gdef\@journalshort{Int. J. Plant Biol.}\gdef\@doiabbr{ijpb}\gdef\@ISSN{2037-0164}\gdef\@continuouspages{yes}}
\DeclareOption{ijt}{\gdef\@journal{ijt}\gdef\@journalfull{International Journal of Topology}\gdef\@journalshort{Int. J. Topol.}\gdef\@doiabbr{ijt}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{ijtm}{\gdef\@journal{ijtm}\gdef\@journalfull{International Journal of Translational Medicine}\gdef\@journalshort{Int. J. Transl. Med.}\gdef\@doiabbr{ijtm}\gdef\@ISSN{2673-8937}\gdef\@continuouspages{yes}}
\DeclareOption{ijtpp}{\gdef\@journal{ijtpp}\gdef\@journalfull{International Journal of Turbomachinery, Propulsion and Power}\gdef\@journalshort{Int. J. Turbomach. Propuls. Power}\gdef\@doiabbr{ijtpp}\gdef\@ISSN{2504-186X}}
\DeclareOption{ime}{\gdef\@journal{ime}\gdef\@journalfull{International Medical Education}\gdef\@journalshort{Int. Med. Educ.}\gdef\@doiabbr{ime}\gdef\@ISSN{2813-141X}\gdef\@continuouspages{yes}}
\DeclareOption{immuno}{\gdef\@journal{immuno}\gdef\@journalfull{Immuno}\gdef\@journalshort{Immuno}\gdef\@doiabbr{immuno}\gdef\@ISSN{2673-5601}\gdef\@continuouspages{yes}}
\DeclareOption{informatics}{\gdef\@journal{informatics}\gdef\@journalfull{Informatics}\gdef\@journalshort{Informatics}\gdef\@doiabbr{informatics}\gdef\@ISSN{2227-9709}}
\DeclareOption{information}{\gdef\@journal{information}\gdef\@journalfull{Information}\gdef\@journalshort{Information}\gdef\@doiabbr{info}\gdef\@ISSN{2078-2489}}
\DeclareOption{infrastructures}{\gdef\@journal{infrastructures}\gdef\@journalfull{Infrastructures}\gdef\@journalshort{Infrastructures}\gdef\@doiabbr{infrastructures}\gdef\@ISSN{2412-3811}}
\DeclareOption{inorganics}{\gdef\@journal{inorganics}\gdef\@journalfull{Inorganics}\gdef\@journalshort{Inorganics}\gdef\@doiabbr{inorganics}\gdef\@ISSN{2304-6740}}
\DeclareOption{insects}{\gdef\@journal{insects}\gdef\@journalfull{Insects}\gdef\@journalshort{Insects}\gdef\@doiabbr{insects}\gdef\@ISSN{2075-4450}}
\DeclareOption{instruments}{\gdef\@journal{instruments}\gdef\@journalfull{Instruments}\gdef\@journalshort{Instruments}\gdef\@doiabbr{instruments}\gdef\@ISSN{2410-390X}}
\DeclareOption{inventions}{\gdef\@journal{inventions}\gdef\@journalfull{Inventions}\gdef\@journalshort{Inventions}\gdef\@doiabbr{inventions}\gdef\@ISSN{2411-5134}}
\DeclareOption{iot}{\gdef\@journal{iot}\gdef\@journalfull{IoT}\gdef\@journalshort{IoT}\gdef\@doiabbr{iot}\gdef\@ISSN{2624-831X}\gdef\@continuouspages{yes}}
\DeclareOption{j}{\gdef\@journal{j}\gdef\@journalfull{J}\gdef\@journalshort{J}\gdef\@doiabbr{j}\gdef\@ISSN{2571-8800}\gdef\@continuouspages{yes}}
\DeclareOption{jal}{\gdef\@journal{jal}\gdef\@journalfull{Journal of Ageing and Longevity}\gdef\@journalshort{J. Ageing Longev.}\gdef\@doiabbr{jal}\gdef\@ISSN{2673-9259}\gdef\@continuouspages{yes}}
\DeclareOption{jcdd}{\gdef\@journal{jcdd}\gdef\@journalfull{Journal of Cardiovascular Development and Disease}\gdef\@journalshort{J. Cardiovasc. Dev. Dis.}\gdef\@doiabbr{jcdd}\gdef\@ISSN{2308-3425}}
\DeclareOption{jcm}{\gdef\@journal{jcm}\gdef\@journalfull{Journal of Clinical Medicine}\gdef\@journalshort{J. Clin. Med.}\gdef\@doiabbr{jcm}\gdef\@ISSN{2077-0383}}
\DeclareOption{jcp}{\gdef\@journal{jcp}\gdef\@journalfull{Journal of Cybersecurity and Privacy}\gdef\@journalshort{J. Cybersecur. Priv.}\gdef\@doiabbr{jcp}\gdef\@ISSN{2624-800X}\gdef\@continuouspages{yes}}
\DeclareOption{jcs}{\gdef\@journal{jcs}\gdef\@journalfull{Journal of Composites Science}\gdef\@journalshort{J. Compos. Sci.}\gdef\@doiabbr{jcs}\gdef\@ISSN{2504-477X}}
\DeclareOption{jcto}{\gdef\@journal{jcto}\gdef\@journalfull{Journal of Clinical \& Translational Ophthalmology}\gdef\@journalshort{J. Clin. Transl. Ophthalmol.}\gdef\@doiabbr{jcto}\gdef\@ISSN{2813-1053}\gdef\@continuouspages{yes}}
\DeclareOption{jdad}{\gdef\@journal{jdad}\gdef\@journalfull{Journal of Dementia and Alzheimer's Disease}\gdef\@journalshort{J. Dement. Alzheimer's Dis.}\gdef\@doiabbr{jdad}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{jdb}{\gdef\@journal{jdb}\gdef\@journalfull{Journal of Developmental Biology}\gdef\@journalshort{J. Dev. Biol.}\gdef\@doiabbr{jdb}\gdef\@ISSN{2221-3759}}
\DeclareOption{jeta}{\gdef\@journal{jeta}\gdef\@journalfull{Journal of Experimental and Theoretical Analyses}\gdef\@journalshort{J. Exp. Theor. Anal.}\gdef\@doiabbr{jeta}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{jfb}{\gdef\@journal{jfb}\gdef\@journalfull{Journal of Functional Biomaterials}\gdef\@journalshort{J. Funct. Biomater.}\gdef\@doiabbr{jfb}\gdef\@ISSN{2079-4983}}
\DeclareOption{jfmk}{\gdef\@journal{jfmk}\gdef\@journalfull{Journal of Functional Morphology and Kinesiology}\gdef\@journalshort{J. Funct. Morphol. Kinesiol.}\gdef\@doiabbr{jfmk}\gdef\@ISSN{2411-5142}}
\DeclareOption{jimaging}{\gdef\@journal{jimaging}\gdef\@journalfull{Journal of Imaging}\gdef\@journalshort{J. Imaging}\gdef\@doiabbr{jimaging}\gdef\@ISSN{2313-433X}}
\DeclareOption{jintelligence}{\gdef\@journal{jintelligence}\gdef\@journalfull{Journal of Intelligence}\gdef\@journalshort{J. Intell.}\gdef\@doiabbr{jintelligence}\gdef\@ISSN{2079-3200}}
\DeclareOption{jlpea}{\gdef\@journal{jlpea}\gdef\@journalfull{Journal of Low Power Electronics and Applications}\gdef\@journalshort{J. Low Power Electron. Appl.}\gdef\@doiabbr{jlpea}\gdef\@ISSN{2079-9268}}
\DeclareOption{jmahp}{\gdef\@journal{jmahp}\gdef\@journalfull{Journal of Market Access \& Health Policy}\gdef\@journalshort{	. Mark. Access Health Policy}\gdef\@doiabbr{jmahp}\gdef\@ISSN{2001-6689}\gdef\@continuouspages{yes}}
\DeclareOption{jmmp}{\gdef\@journal{jmmp}\gdef\@journalfull{Journal of Manufacturing and Materials Processing}\gdef\@journalshort{J. Manuf. Mater. Process.}\gdef\@doiabbr{jmmp}\gdef\@ISSN{2504-4494}}
\DeclareOption{jmp}{\gdef\@journal{jmp}\gdef\@journalfull{Journal of Molecular Pathology}\gdef\@journalshort{J. Mol. Pathol.}\gdef\@doiabbr{jmp}\gdef\@ISSN{2673-5261}\gdef\@continuouspages{yes}}
\DeclareOption{jmse}{\gdef\@journal{jmse}\gdef\@journalfull{Journal of Marine Science and Engineering}\gdef\@journalshort{J. Mar. Sci. Eng.}\gdef\@doiabbr{jmse}\gdef\@ISSN{2077-1312}}
\DeclareOption{jne}{\gdef\@journal{jne}\gdef\@journalfull{Journal of Nuclear Engineering}\gdef\@journalshort{J. Nucl. Eng.}\gdef\@doiabbr{jne}\gdef\@ISSN{2673-4362}\gdef\@continuouspages{yes}}
\DeclareOption{jnt}{\gdef\@journal{jnt}\gdef\@journalfull{Journal of Nanotheranostics}\gdef\@journalshort{J. Nanotheranostics}\gdef\@doiabbr{jnt}\gdef\@ISSN{2624-845X}\gdef\@continuouspages{yes}}
\DeclareOption{jof}{\gdef\@journal{jof}\gdef\@journalfull{Journal of Fungi}\gdef\@journalshort{J. Fungi}\gdef\@doiabbr{jof}\gdef\@ISSN{2309-608X}}
\DeclareOption{joitmc}{\gdef\@journal{joitmc}\gdef\@journalfull{Journal of Open Innovation: Technology, Market, and Complexity}\gdef\@journalshort{J. Open Innov. Technol. Mark. Complex.}\gdef\@doiabbr{joitmc}\gdef\@ISSN{2199-8531}}
\DeclareOption{jor}{\gdef\@journal{jor}\gdef\@journalfull{Journal of Respiration}\gdef\@journalshort{J. Respir.}\gdef\@doiabbr{jor}\gdef\@ISSN{2673-527X}\gdef\@continuouspages{yes}}
\DeclareOption{journalmedia}{\gdef\@journal{journalmedia}\gdef\@journalfull{Journalism and Media}\gdef\@journalshort{Journal. Media}\gdef\@doiabbr{journalmedia}\gdef\@ISSN{2673-5172}\gdef\@continuouspages{yes}}
\DeclareOption{jox}{\gdef\@journal{jox}\gdef\@journalfull{Journal of Xenobiotics}\gdef\@journalshort{J. Xenobiot.}\gdef\@doiabbr{jox}\gdef\@ISSN{2039-4713}\gdef\@continuouspages{yes}}
\DeclareOption{jpm}{\gdef\@journal{jpm}\gdef\@journalfull{Journal of Personalized Medicine}\gdef\@journalshort{J. Pers. Med.}\gdef\@doiabbr{jpm}\gdef\@ISSN{2075-4426}}
\DeclareOption{jrfm}{\gdef\@journal{jrfm}\gdef\@journalfull{Journal of Risk and Financial Management}\gdef\@journalshort{J. Risk Financial Manag.}\gdef\@doiabbr{jrfm}\gdef\@ISSN{1911-8074}}
\DeclareOption{jsan}{\gdef\@journal{jsan}\gdef\@journalfull{Journal of Sensor and Actuator Networks}\gdef\@journalshort{J. Sens. Actuator Netw.}\gdef\@doiabbr{jsan}\gdef\@ISSN{2224-2708}}
\DeclareOption{jtaer}{\gdef\@journal{jtaer}\gdef\@journalfull{Journal of Theoretical and Applied Electronic Commerce Research}\gdef\@journalshort{J. Theor. Appl. Electron. Commer. Res.}\gdef\@doiabbr{jtaer}\gdef\@ISSN{0718-1876}\gdef\@continuouspages{yes}}
\DeclareOption{jvd}{\gdef\@journal{jvd}\gdef\@journalfull{Journal of Vascular Diseases}\gdef\@journalshort{JVD}\gdef\@doiabbr{jvd}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{jzbg}{\gdef\@journal{jzbg}\gdef\@journalfull{Journal of Zoological and Botanical Gardens}\gdef\@journalshort{J. Zool. Bot. Gard.}\gdef\@doiabbr{jzbg}\gdef\@ISSN{2673-5636}\gdef\@continuouspages{yes}}
%\DeclareOption{kidney}{\gdef\@journal{kidney}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{kidneydial}{\gdef\@journal{kidneydial}\gdef\@journalfull{Kidney and Dialysis}\gdef\@journalshort{Kidney Dial.}\gdef\@doiabbr{kinasesphosphatases}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{kinasesphosphatases}{\gdef\@journal{kinasesphosphatases}\gdef\@journalfull{Kinases and Phosphatases}\gdef\@journalshort{Kinases Phosphatases}\gdef\@doiabbr{kinasesphosphatases}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{knowledge}{\gdef\@journal{knowledge}\gdef\@journalfull{Knowledge}\gdef\@journalshort{Knowledge}\gdef\@doiabbr{knowledge}\gdef\@ISSN{2673-9585}\gdef\@continuouspages{yes}}
\DeclareOption{land}{\gdef\@journal{land}\gdef\@journalfull{Land}\gdef\@journalshort{Land}\gdef\@doiabbr{land}\gdef\@ISSN{2073-445X}}
\DeclareOption{languages}{\gdef\@journal{languages}\gdef\@journalfull{Languages}\gdef\@journalshort{Languages}\gdef\@doiabbr{languages}\gdef\@ISSN{2226-471X}}
\DeclareOption{laws}{\gdef\@journal{laws}\gdef\@journalfull{Laws}\gdef\@journalshort{Laws}\gdef\@doiabbr{laws}\gdef\@ISSN{2075-471X}}
\DeclareOption{life}{\gdef\@journal{life}\gdef\@journalfull{Life}\gdef\@journalshort{Life}\gdef\@doiabbr{life}\gdef\@ISSN{2075-1729}}
\DeclareOption{limnolrev}{\gdef\@journal{limnolrev}\gdef\@journalfull{Limnological Review}\gdef\@journalshort{Limnol. Rev.}\gdef\@doiabbr{limnolrev}\gdef\@ISSN{2300-7575}\gdef\@continuouspages{yes}}
\DeclareOption{lipidology}{\gdef\@journal{lipidology}\gdef\@journalfull{Lipidology}\gdef\@journalshort{Lipidology}\gdef\@doiabbr{lipidology}\gdef\@ISSN{2813-7086}\gdef\@continuouspages{yes}}
\DeclareOption{liquids}{\gdef\@journal{liquids}\gdef\@journalfull{Liquids}\gdef\@journalshort{Liquids}\gdef\@doiabbr{liquids}\gdef\@ISSN{2673-8015}\gdef\@continuouspages{yes}}
\DeclareOption{literature}{\gdef\@journal{literature}\gdef\@journalfull{Literature}\gdef\@journalshort{Literature}\gdef\@doiabbr{literature}\gdef\@ISSN{2410-9789}\gdef\@continuouspages{yes}}
\DeclareOption{livers}{\gdef\@journal{livers}\gdef\@journalfull{Livers}\gdef\@journalshort{Livers}\gdef\@doiabbr{livers}\gdef\@ISSN{2673-4389}\gdef\@continuouspages{yes}}
\DeclareOption{logics}{\gdef\@journal{logics}\gdef\@journalfull{Logics}\gdef\@journalshort{Logics}\gdef\@doiabbr{logics}\gdef\@ISSN{2813-0405}\gdef\@continuouspages{yes}}
\DeclareOption{logistics}{\gdef\@journal{logistics}\gdef\@journalfull{Logistics}\gdef\@journalshort{Logistics}\gdef\@doiabbr{logistics}\gdef\@ISSN{2305-6290}}
\DeclareOption{lubricants}{\gdef\@journal{lubricants}\gdef\@journalfull{Lubricants}\gdef\@journalshort{Lubricants}\gdef\@doiabbr{lubricants}\gdef\@ISSN{2075-4442}}
\DeclareOption{lymphatics}{\gdef\@journal{lymphatics}\gdef\@journalfull{Lymphatics}\gdef\@journalshort{Lymphatics}\gdef\@doiabbr{lymphatics}\gdef\@ISSN{}\gdef\@continuouspages{yes}}	
\DeclareOption{machines}{\gdef\@journal{machines}\gdef\@journalfull{Machines}\gdef\@journalshort{Machines}\gdef\@doiabbr{machines}\gdef\@ISSN{2075-1702}}
\DeclareOption{macromol}{\gdef\@journal{macromol}\gdef\@journalfull{Macromol}\gdef\@journalshort{Macromol}\gdef\@doiabbr{macromol}\gdef\@ISSN{2673-6209}\gdef\@continuouspages{yes}}
\DeclareOption{magnetism}{\gdef\@journal{magnetism}\gdef\@journalfull{Magnetism}\gdef\@journalshort{Magnetism}\gdef\@doiabbr{magnetism}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{magnetochemistry}{\gdef\@journal{magnetochemistry}\gdef\@journalfull{Magnetochemistry}\gdef\@journalshort{Magnetochemistry}\gdef\@doiabbr{magnetochemistry}\gdef\@ISSN{2312-7481}}
\DeclareOption{make}{\gdef\@journal{make}\gdef\@journalfull{Machine Learning and Knowledge Extraction}\gdef\@journalshort{Mach. Learn. Knowl. Extr.}\gdef\@doiabbr{make}\gdef\@ISSN{2504-4990}\gdef\@continuouspages{yes}}
\DeclareOption{marinedrugs}{\gdef\@journal{marinedrugs}\gdef\@journalfull{Marine Drugs}\gdef\@journalshort{Mar. Drugs}\gdef\@doiabbr{md}\gdef\@ISSN{1660-3397}}
\DeclareOption{materials}{\gdef\@journal{materials}\gdef\@journalfull{Materials}\gdef\@journalshort{Materials}\gdef\@doiabbr{ma}\gdef\@ISSN{1996-1944}}
\DeclareOption{materproc}{\gdef\@journal{materproc}\gdef\@journalfull{Materials Proceedings}\gdef\@journalshort{Mater. Proc.}\gdef\@doiabbr{materproc}\gdef\@ISSN{2673-4605}}
\DeclareOption{mathematics}{\gdef\@journal{mathematics}\gdef\@journalfull{Mathematics}\gdef\@journalshort{Mathematics}\gdef\@doiabbr{math}\gdef\@ISSN{2227-7390}}
\DeclareOption{mca}{\gdef\@journal{mca}\gdef\@journalfull{Mathematical and Computational Applications}\gdef\@journalshort{Math. Comput. Appl.}\gdef\@doiabbr{mca}\gdef\@ISSN{2297-8747}}
\DeclareOption{measurements}{\gdef\@journal{measurements}\gdef\@journalfull{Measurements}\gdef\@journalshort{Measurements}\gdef\@doiabbr{measurements}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{medicina}{\gdef\@journal{medicina}\gdef\@journalfull{Medicina}\gdef\@journalshort{Medicina}\gdef\@doiabbr{medicina}\gdef\@ISSN{1648-9144}}
\DeclareOption{medicines}{\gdef\@journal{medicines}\gdef\@journalfull{Medicines}\gdef\@journalshort{Medicines}\gdef\@doiabbr{medicines}\gdef\@ISSN{2305-6320}}
\DeclareOption{medsci}{\gdef\@journal{medsci}\gdef\@journalfull{Medical Sciences}\gdef\@journalshort{Med. Sci.}\gdef\@doiabbr{medsci}\gdef\@ISSN{2076-3271}}
\DeclareOption{membranes}{\gdef\@journal{membranes}\gdef\@journalfull{Membranes}\gdef\@journalshort{Membranes}\gdef\@doiabbr{membranes}\gdef\@ISSN{2077-0375}}
\DeclareOption{merits}{\gdef\@journal{merits}\gdef\@journalfull{Merits}\gdef\@journalshort{Merits}\gdef\@doiabbr{merits}\gdef\@ISSN{2673-8104}\gdef\@continuouspages{yes}}
\DeclareOption{metabolites}{\gdef\@journal{metabolites}\gdef\@journalfull{Metabolites}\gdef\@journalshort{Metabolites}\gdef\@doiabbr{metabo}\gdef\@ISSN{2218-1989}}
\DeclareOption{metals}{\gdef\@journal{metals}\gdef\@journalfull{Metals}\gdef\@journalshort{Metals}\gdef\@doiabbr{met}\gdef\@ISSN{2075-4701}}
\DeclareOption{meteorology}{\gdef\@journal{meteorology}\gdef\@journalfull{Meteorology}\gdef\@journalshort{Meteorology}\gdef\@doiabbr{meteorology}\gdef\@ISSN{2674-0494}\gdef\@continuouspages{yes}}
\DeclareOption{methane}{\gdef\@journal{methane}\gdef\@journalfull{Methane}\gdef\@journalshort{Methane}\gdef\@doiabbr{methane}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{metrology}{\gdef\@journal{metrology}\gdef\@journalfull{Metrology}\gdef\@journalshort{Metrology}\gdef\@doiabbr{metrology}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{micro}{\gdef\@journal{micro}\gdef\@journalfull{Micro}\gdef\@journalshort{Micro}\gdef\@doiabbr{micro}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{microarrays}{\gdef\@journal{microarrays}\gdef\@journalfull{Microarrays}\gdef\@journalshort{Microarrays}\gdef\@doiabbr{microarrays}\gdef\@ISSN{2076-3905}}
\DeclareOption{microbiolres}{\gdef\@journal{microbiolres}\gdef\@journalfull{Microbiology Research}\gdef\@journalshort{Microbiol. Res.}\gdef\@doiabbr{microbiolres}\gdef\@ISSN{2036-7481}\gdef\@continuouspages{yes}}
\DeclareOption{micromachines}{\gdef\@journal{micromachines}\gdef\@journalfull{Micromachines}\gdef\@journalshort{Micromachines}\gdef\@doiabbr{mi}\gdef\@ISSN{2072-666X}}
\DeclareOption{microorganisms}{\gdef\@journal{microorganisms}\gdef\@journalfull{Microorganisms}\gdef\@journalshort{Microorganisms}\gdef\@doiabbr{microorganisms}\gdef\@ISSN{2076-2607}}
\DeclareOption{microplastics}{\gdef\@journal{microplastics}\gdef\@journalfull{Microplastics}\gdef\@journalshort{Microplastics}\gdef\@doiabbr{microplastics}\gdef\@ISSN{2673-8929}\gdef\@continuouspages{yes}}
\DeclareOption{minerals}{\gdef\@journal{minerals}\gdef\@journalfull{Minerals}\gdef\@journalshort{Minerals}\gdef\@doiabbr{min}\gdef\@ISSN{2075-163X}}
\DeclareOption{mining}{\gdef\@journal{mining}\gdef\@journalfull{Mining}\gdef\@journalshort{Mining}\gdef\@doiabbr{mining}\gdef\@ISSN{2673-6489}\gdef\@continuouspages{yes}}
\DeclareOption{modelling}{\gdef\@journal{modelling}\gdef\@journalfull{Modelling}\gdef\@journalshort{Modelling}\gdef\@doiabbr{modelling}\gdef\@ISSN{2673-3951}\gdef\@continuouspages{yes}}
\DeclareOption{molbank}{\gdef\@journal{molbank}\gdef\@journalfull{Molbank}\gdef\@journalshort{Molbank}\gdef\@doiabbr{M}\gdef\@ISSN{1422-8599}}
\DeclareOption{molecules}{\gdef\@journal{molecules}\gdef\@journalfull{Molecules}\gdef\@journalshort{Molecules}\gdef\@doiabbr{molecules}\gdef\@ISSN{1420-3049}}
\DeclareOption{mps}{\gdef\@journal{mps}\gdef\@journalfull{Methods and Protocols}\gdef\@journalshort{Methods Protoc.}\gdef\@doiabbr{mps}\gdef\@ISSN{2409-9279}}
\DeclareOption{msf}{\gdef\@journal{msf}\gdef\@journalfull{Medical Sciences Forum}\gdef\@journalshort{Med. Sci. Forum}\gdef\@doiabbr{msf}\gdef\@ISSN{2673-9992}}
\DeclareOption{mti}{\gdef\@journal{mti}\gdef\@journalfull{Multimodal Technologies and Interaction}\gdef\@journalshort{Multimodal Technol. Interact.}\gdef\@doiabbr{mti}\gdef\@ISSN{2414-4088}}
\DeclareOption{muscles}{\gdef\@journal{muscles}\gdef\@journalfull{Muscles}\gdef\@journalshort{Muscles}\gdef\@doiabbr{muscles}\gdef\@ISSN{2813-0413}\gdef\@continuouspages{yes}}
\DeclareOption{nanoenergyadv}{\gdef\@journal{nanoenergyadv}\gdef\@journalfull{Nanoenergy Advances}\gdef\@journalshort{Nanoenergy Adv.}\gdef\@doiabbr{nanoenergyadv}\gdef\@ISSN{2673-706X}\gdef\@continuouspages{yes}}
\DeclareOption{nanomanufacturing}{\gdef\@journal{nanomanufacturing}\gdef\@journalfull{Nanomanufacturing}\gdef\@journalshort{Nanomanufacturing}\gdef\@doiabbr{nanomanufacturing}\gdef\@ISSN{2673-687X}\gdef\@continuouspages{yes}}
\DeclareOption{nanomaterials}{\gdef\@journal{nanomaterials}\gdef\@journalfull{Nanomaterials}\gdef\@journalshort{Nanomaterials}\gdef\@doiabbr{nano}\gdef\@ISSN{2079-4991}}
\DeclareOption{ncrna}{\gdef\@journal{ncrna}\gdef\@journalfull{Non-Coding RNA}\gdef\@journalshort{Non-Coding RNA}\gdef\@doiabbr{ncrna}\gdef\@ISSN{2311-553X}}
\DeclareOption{ndt}{\gdef\@journal{ndt}\gdef\@journalfull{NDT}\gdef\@journalshort{NDT}\gdef\@doiabbr{ndt}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{network}{\gdef\@journal{network}\gdef\@journalfull{Network}\gdef\@journalshort{Network}\gdef\@doiabbr{network}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{neuroglia}{\gdef\@journal{neuroglia}\gdef\@journalfull{Neuroglia}\gdef\@journalshort{Neuroglia}\gdef\@doiabbr{neuroglia}\gdef\@ISSN{2571-6980}\gdef\@continuouspages{yes}}
\DeclareOption{neurolint}{\gdef\@journal{neurolint}\gdef\@journalfull{Neurology International}\gdef\@journalshort{Neurol. Int.}\gdef\@doiabbr{neurolint}\gdef\@ISSN{2035-8377}\gdef\@continuouspages{yes}}
\DeclareOption{neurosci}{\gdef\@journal{neurosci}\gdef\@journalfull{NeuroSci}\gdef\@journalshort{NeuroSci}\gdef\@doiabbr{neurosci}\gdef\@ISSN{2673-4087}\gdef\@continuouspages{yes}}
\DeclareOption{nitrogen}{\gdef\@journal{nitrogen}\gdef\@journalfull{Nitrogen}\gdef\@journalshort{Nitrogen}\gdef\@doiabbr{nitrogen}\gdef\@ISSN{2504-3129}\gdef\@continuouspages{yes}}
\DeclareOption{notspecified}{\gdef\@journal{notspecified}\gdef\@journalfull{Journal Not Specified}\gdef\@journalshort{Journal Not Specified}\gdef\@doiabbr{}\gdef\@ISSN{}}
%%\DeclareOption{nri}{\gdef\@journal{nri}\gdef\@journalfull{Natural Resources Informatics}\gdef\@journalshort{Nat. Resour. Inform.}\gdef\@doiabbr{nri}\gdef\@ISSN{2673-981X}\gdef\@continuouspages{yes}}
\DeclareOption{nursrep}{\gdef\@journal{nursrep}\gdef\@journalfull{Nursing Reports}\gdef\@journalshort{Nurs. Rep.}\gdef\@doiabbr{nursrep}\gdef\@ISSN{2039-4403}\gdef\@continuouspages{yes}}
\DeclareOption{nutraceuticals}{\gdef\@journal{nutraceuticals}\gdef\@journalfull{Nutraceuticals}\gdef\@journalshort{Nutraceuticals}\gdef\@doiabbr{nutraceuticals}\gdef\@ISSN{1661-3821}\gdef\@continuouspages{yes}}
\DeclareOption{nutrients}{\gdef\@journal{nutrients}\gdef\@journalfull{Nutrients}\gdef\@journalshort{Nutrients}\gdef\@doiabbr{nu}\gdef\@ISSN{2072-6643}}
\DeclareOption{obesities}{\gdef\@journal{obesities}\gdef\@journalfull{Obesities}\gdef\@journalshort{Obesities}\gdef\@doiabbr{obesities}\gdef\@ISSN{2673-4168}\gdef\@continuouspages{yes}}
\DeclareOption{oceans}{\gdef\@journal{oceans}\gdef\@journalfull{Oceans}\gdef\@journalshort{Oceans}\gdef\@doiabbr{oceans}\gdef\@ISSN{2673-1924}\gdef\@continuouspages{yes}}
\DeclareOption{ohbm}{\gdef\@journal{ohbm}\gdef\@journalfull{Journal of Otorhinolaryngology, Hearing and Balance Medicine}\gdef\@journalshort{J. Otorhinolaryngol. Hear. Balance Med.}\gdef\@doiabbr{ohbm}\gdef\@ISSN{2504-463X}}
\DeclareOption{onco}{\gdef\@journal{onco}\gdef\@journalfull{Onco}\gdef\@journalshort{Onco}\gdef\@doiabbr{onco}\gdef\@ISSN{2673-7523}\gdef\@continuouspages{yes}}
%\DeclareOption{oncopathology}{\gdef\@journal{oncopathology}\gdef\@journalfull{}\gdef\@journalshort{}\gdef\@doiabbr{}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{optics}{\gdef\@journal{optics}\gdef\@journalfull{Optics}\gdef\@journalshort{Optics}\gdef\@doiabbr{opt}\gdef\@ISSN{2673-3269}\gdef\@continuouspages{yes}}
\DeclareOption{oral}{\gdef\@journal{oral}\gdef\@journalfull{Oral}\gdef\@journalshort{Oral}\gdef\@doiabbr{oral}\gdef\@ISSN{2673-6373}\gdef\@continuouspages{yes}}
\DeclareOption{organics}{\gdef\@journal{organics}\gdef\@journalfull{Organics}\gdef\@journalshort{Organics}\gdef\@doiabbr{org}\gdef\@ISSN{2673-401X}\gdef\@continuouspages{yes}}
\DeclareOption{organoids}{\gdef\@journal{organoids}\gdef\@journalfull{Organoids}\gdef\@journalshort{Organoids}\gdef\@doiabbr{organoids}\gdef\@ISSN{2674-1172}\gdef\@continuouspages{yes}}
\DeclareOption{osteology}{\gdef\@journal{osteology}\gdef\@journalfull{Osteology}\gdef\@journalshort{Osteology}\gdef\@doiabbr{osteology}\gdef\@ISSN{2673-4036}\gdef\@continuouspages{yes}}
\DeclareOption{oxygen}{\gdef\@journal{oxygen}\gdef\@journalfull{Oxygen}\gdef\@journalshort{Oxygen}\gdef\@doiabbr{oxygen}\gdef\@ISSN{2673-9801}\gdef\@continuouspages{yes}}
\DeclareOption{parasites}{\gdef\@journal{parasites}\gdef\@journalfull{Parasites}\gdef\@journalshort{Parasites}\gdef\@doiabbr{parasites}\gdef\@ISSN{2673-5946}}
\DeclareOption{parasitologia}{\gdef\@journal{parasitologia}\gdef\@journalfull{Parasitologia}\gdef\@journalshort{Parasitologia}\gdef\@doiabbr{parasitologia}\gdef\@ISSN{2673-6772}\gdef\@continuouspages{yes}}
\DeclareOption{particles}{\gdef\@journal{particles}\gdef\@journalfull{Particles}\gdef\@journalshort{Particles}\gdef\@doiabbr{particles}\gdef\@ISSN{2571-712X}\gdef\@continuouspages{yes}}
\DeclareOption{pathogens}{\gdef\@journal{pathogens}\gdef\@journalfull{Pathogens}\gdef\@journalshort{Pathogens}\gdef\@doiabbr{pathogens}\gdef\@ISSN{2076-0817}}
\DeclareOption{pathophysiology}{\gdef\@journal{pathophysiology}\gdef\@journalfull{Pathophysiology}\gdef\@journalshort{Pathophysiology}\gdef\@doiabbr{pathophysiology}\gdef\@ISSN{0928-4680}\gdef\@continuouspages{yes}}
\DeclareOption{pediatrrep}{\gdef\@journal{pediatrrep}\gdef\@journalfull{Pediatric Reports}\gdef\@journalshort{Pediatr. Rep.}\gdef\@doiabbr{pediatric}\gdef\@ISSN{2036-7503}\gdef\@continuouspages{yes}}
\DeclareOption{pets}{\gdef\@journal{pets}\gdef\@journalfull{Pets}\gdef\@journalshort{Pets}\gdef\@doiabbr{pets}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{pharmaceuticals}{\gdef\@journal{pharmaceuticals}\gdef\@journalfull{Pharmaceuticals}\gdef\@journalshort{Pharmaceuticals}\gdef\@doiabbr{ph}\gdef\@ISSN{1424-8247}}
\DeclareOption{pharmaceutics}{\gdef\@journal{pharmaceutics}\gdef\@journalfull{Pharmaceutics}\gdef\@journalshort{Pharmaceutics}\gdef\@doiabbr{pharmaceutics}\gdef\@ISSN{1999-4923}}
\DeclareOption{pharmacoepidemiology}{\gdef\@journal{pharmacoepidemiology}\gdef\@journalfull{Pharmacoepidemiology}\gdef\@journalshort{Pharmacoepidemiology}\gdef\@doiabbr{pharma}\gdef\@ISSN{2813-0618}\gdef\@continuouspages{yes}}
\DeclareOption{pharmacy}{\gdef\@journal{pharmacy}\gdef\@journalfull{Pharmacy}\gdef\@journalshort{Pharmacy}\gdef\@doiabbr{pharmacy}\gdef\@ISSN{2226-4787}}
\DeclareOption{philosophies}{\gdef\@journal{philosophies}\gdef\@journalfull{Philosophies}\gdef\@journalshort{Philosophies}\gdef\@doiabbr{philosophies}\gdef\@ISSN{2409-9287}}
\DeclareOption{photochem}{\gdef\@journal{photochem}\gdef\@journalfull{Photochem}\gdef\@journalshort{Photochem}\gdef\@doiabbr{photochem}\gdef\@ISSN{2673-7256}\gdef\@continuouspages{yes}}
\DeclareOption{photonics}{\gdef\@journal{photonics}\gdef\@journalfull{Photonics}\gdef\@journalshort{Photonics}\gdef\@doiabbr{photonics}\gdef\@ISSN{2304-6732}}
\DeclareOption{phycology}{\gdef\@journal{phycology}\gdef\@journalfull{Phycology}\gdef\@journalshort{Phycology}\gdef\@doiabbr{phycology}\gdef\@ISSN{2673-9410}\gdef\@continuouspages{yes}}
\DeclareOption{physchem}{\gdef\@journal{physchem}\gdef\@journalfull{Physchem}\gdef\@journalshort{Physchem}\gdef\@doiabbr{physchem}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{physics}{\gdef\@journal{physics}\gdef\@journalfull{Physics}\gdef\@journalshort{Physics}\gdef\@doiabbr{physics}\gdef\@ISSN{2624-8174}\gdef\@continuouspages{yes}}
\DeclareOption{physiologia}{\gdef\@journal{physiologia}\gdef\@journalfull{Physiologia}\gdef\@journalshort{Physiologia}\gdef\@doiabbr{physiologia}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{plants}{\gdef\@journal{plants}\gdef\@journalfull{Plants}\gdef\@journalshort{Plants}\gdef\@doiabbr{plants}\gdef\@ISSN{2223-7747}}
\DeclareOption{plasma}{\gdef\@journal{plasma}\gdef\@journalfull{Plasma}\gdef\@journalshort{Plasma}\gdef\@doiabbr{plasma}\gdef\@ISSN{2571-6182}\gdef\@continuouspages{yes}}
\DeclareOption{platforms}{\gdef\@journal{platforms}\gdef\@journalfull{Platforms}\gdef\@journalshort{Platforms}\gdef\@doiabbr{platforms}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{pollutants}{\gdef\@journal{pollutants}\gdef\@journalfull{Pollutants}\gdef\@journalshort{Pollutants}\gdef\@doiabbr{pollutants}\gdef\@ISSN{2673-4672}\gdef\@continuouspages{yes}}
\DeclareOption{polymers}{\gdef\@journal{polymers}\gdef\@journalfull{Polymers}\gdef\@journalshort{Polymers}\gdef\@doiabbr{polym}\gdef\@ISSN{2073-4360}}
\DeclareOption{polysaccharides}{\gdef\@journal{polysaccharides}\gdef\@journalfull{Polysaccharides}\gdef\@journalshort{Polysaccharides}\gdef\@doiabbr{polysaccharides}\gdef\@ISSN{2673-4176}\gdef\@continuouspages{yes}}
\DeclareOption{poultry}{\gdef\@journal{poultry}\gdef\@journalfull{Poultry}\gdef\@journalshort{Poultry}\gdef\@doiabbr{poultry}\gdef\@ISSN{2674-1164}\gdef\@continuouspages{yes}}
\DeclareOption{powders}{\gdef\@journal{powders}\gdef\@journalfull{Powders}\gdef\@journalshort{Powders}\gdef\@doiabbr{powders}\gdef\@ISSN{2674-0516}\gdef\@continuouspages{yes}}
\DeclareOption{preprints}{\gdef\@journal{preprints}\gdef\@journalfull{Preprints}\gdef\@journalshort{Preprints}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{proceedings}{\gdef\@journal{proceedings}\gdef\@journalfull{Proceedings}\gdef\@journalshort{Proceedings}\gdef\@doiabbr{proceedings}\gdef\@ISSN{2504-3900}}
\DeclareOption{processes}{\gdef\@journal{processes}\gdef\@journalfull{Processes}\gdef\@journalshort{Processes}\gdef\@doiabbr{pr}\gdef\@ISSN{2227-9717}}
\DeclareOption{prosthesis}{\gdef\@journal{prosthesis}\gdef\@journalfull{Prosthesis}\gdef\@journalshort{Prosthesis}\gdef\@doiabbr{prosthesis}\gdef\@ISSN{2673-1592}\gdef\@continuouspages{yes}}
\DeclareOption{proteomes}{\gdef\@journal{proteomes}\gdef\@journalfull{Proteomes}\gdef\@journalshort{Proteomes}\gdef\@doiabbr{proteomes}\gdef\@ISSN{2227-7382}}
\DeclareOption{psf}{\gdef\@journal{psf}\gdef\@journalfull{Physical Sciences Forum}\gdef\@journalshort{Phys. Sci. Forum}\gdef\@doiabbr{psf}\gdef\@ISSN{2673-9984}}
\DeclareOption{psych}{\gdef\@journal{psych}\gdef\@journalfull{Psych}\gdef\@journalshort{Psych}\gdef\@doiabbr{psych}\gdef\@ISSN{2624-8611}\gdef\@continuouspages{yes}}
\DeclareOption{psychiatryint}{\gdef\@journal{psychiatryint}\gdef\@journalfull{Psychiatry International}\gdef\@journalshort{Psychiatry Int.}\gdef\@doiabbr{psychiatryint}\gdef\@ISSN{2673-5318}\gdef\@continuouspages{yes}}
\DeclareOption{psychoactives}{\gdef\@journal{psychoactives}\gdef\@journalfull{Psychoactives}\gdef\@journalshort{Psychoactives}\gdef\@doiabbr{psychoactives}\gdef\@ISSN{2813-1851}\gdef\@continuouspages{yes}}
\DeclareOption{publications}{\gdef\@journal{publications}\gdef\@journalfull{Publications}\gdef\@journalshort{Publications}\gdef\@doiabbr{publications}\gdef\@ISSN{2304-6775}}
\DeclareOption{quantumrep}{\gdef\@journal{quantumrep}\gdef\@journalfull{Quantum Reports}\gdef\@journalshort{Quantum Rep.}\gdef\@doiabbr{quantum}\gdef\@ISSN{2624-960X}\gdef\@continuouspages{yes}}
\DeclareOption{quaternary}{\gdef\@journal{quaternary}\gdef\@journalfull{Quaternary}\gdef\@journalshort{Quaternary}\gdef\@doiabbr{quat}\gdef\@ISSN{2571-550X}}
\DeclareOption{qubs}{\gdef\@journal{qubs}\gdef\@journalfull{Quantum Beam Science}\gdef\@journalshort{Quantum Beam Sci.}\gdef\@doiabbr{qubs}\gdef\@ISSN{2412-382X}}
\DeclareOption{radiation}{\gdef\@journal{radiation}\gdef\@journalfull{Radiation}\gdef\@journalshort{Radiation}\gdef\@doiabbr{radiation}\gdef\@ISSN{2673-592X}\gdef\@continuouspages{yes}}
\DeclareOption{reactions}{\gdef\@journal{reactions}\gdef\@journalfull{Reactions}\gdef\@journalshort{Reactions}\gdef\@doiabbr{reactions}\gdef\@ISSN{2624-781X}\gdef\@continuouspages{yes}}
\DeclareOption{realestate}{\gdef\@journal{realestate}\gdef\@journalfull{Real Estate}\gdef\@journalshort{Real Estate}\gdef\@doiabbr{realestate}\gdef\@ISSN{2813-8090}\gdef\@continuouspages{yes}}
\DeclareOption{receptors}{\gdef\@journal{receptors}\gdef\@journalfull{Receptors}\gdef\@journalshort{Receptors}\gdef\@doiabbr{receptors}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{recycling}{\gdef\@journal{recycling}\gdef\@journalfull{Recycling}\gdef\@journalshort{Recycling}\gdef\@doiabbr{recycling}\gdef\@ISSN{2313-4321}}
\DeclareOption{regeneration}{\gdef\@journal{regeneration}\gdef\@journalfull{Regeneration}\gdef\@journalshort{Regeneration}\gdef\@doiabbr{regeneration}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{religions}{\gdef\@journal{religions}\gdef\@journalfull{Religions}\gdef\@journalshort{Religions}\gdef\@doiabbr{rel}\gdef\@ISSN{2077-1444}}
\DeclareOption{remotesensing}{\gdef\@journal{remotesensing}\gdef\@journalfull{Remote Sensing}\gdef\@journalshort{Remote Sens.}\gdef\@doiabbr{rs}\gdef\@ISSN{2072-4292}}
\DeclareOption{reports}{\gdef\@journal{reports}\gdef\@journalfull{Reports}\gdef\@journalshort{Reports}\gdef\@doiabbr{reports}\gdef\@ISSN{2571-841X}}
\DeclareOption{reprodmed}{\gdef\@journal{reprodmed}\gdef\@journalfull{Reproductive Medicine}\gdef\@journalshort{Reprod. Med.}\gdef\@doiabbr{reprodmed}\gdef\@ISSN{2673-3897}\gdef\@continuouspages{yes}}
\DeclareOption{resources}{\gdef\@journal{resources}\gdef\@journalfull{Resources}\gdef\@journalshort{Resources}\gdef\@doiabbr{resources}\gdef\@ISSN{2079-9276}}
\DeclareOption{rheumato}{\gdef\@journal{rheumato}\gdef\@journalfull{Rheumato}\gdef\@journalshort{Rheumato}\gdef\@doiabbr{rheumato}\gdef\@ISSN{2674-0621}\gdef\@continuouspages{yes}}
\DeclareOption{risks}{\gdef\@journal{risks}\gdef\@journalfull{Risks}\gdef\@journalshort{Risks}\gdef\@doiabbr{risks}\gdef\@ISSN{2227-9091}}
\DeclareOption{robotics}{\gdef\@journal{robotics}\gdef\@journalfull{Robotics}\gdef\@journalshort{Robotics}\gdef\@doiabbr{robotics}\gdef\@ISSN{2218-6581}}
\DeclareOption{ruminants}{\gdef\@journal{ruminants}\gdef\@journalfull{Ruminants}\gdef\@journalshort{Ruminants}\gdef\@doiabbr{ruminants}\gdef\@ISSN{2673-933X}\gdef\@continuouspages{yes}}
\DeclareOption{safety}{\gdef\@journal{safety}\gdef\@journalfull{Safety}\gdef\@journalshort{Safety}\gdef\@doiabbr{safety}\gdef\@ISSN{2313-576X}}
\DeclareOption{sci}{\gdef\@journal{sci}\gdef\@journalfull{Sci}\gdef\@journalshort{Sci}\gdef\@doiabbr{sci}\gdef\@ISSN{2413-4155}}
\DeclareOption{scipharm}{\gdef\@journal{scipharm}\gdef\@journalfull{Scientia Pharmaceutica}\gdef\@journalshort{Sci. Pharm.}\gdef\@doiabbr{scipharm}\gdef\@ISSN{2218-0532}}
\DeclareOption{sclerosis}{\gdef\@journal{sclerosis}\gdef\@journalfull{Sclerosis}\gdef\@journalshort{Sclerosis}\gdef\@doiabbr{sclerosis}\gdef\@ISSN{2813-3064}\gdef\@continuouspages{yes}}
\DeclareOption{seeds}{\gdef\@journal{seeds}\gdef\@journalfull{Seeds}\gdef\@journalshort{Seeds}\gdef\@doiabbr{seeds}\gdef\@ISSN{2674-1024}\gdef\@continuouspages{yes}}
\DeclareOption{sensors}{\gdef\@journal{sensors}\gdef\@journalfull{Sensors}\gdef\@journalshort{Sensors}\gdef\@doiabbr{s}\gdef\@ISSN{1424-8220}}
\DeclareOption{separations}{\gdef\@journal{separations}\gdef\@journalfull{Separations}\gdef\@journalshort{Separations}\gdef\@doiabbr{separations}\gdef\@ISSN{2297-8739}}
\DeclareOption{sexes}{\gdef\@journal{sexes}\gdef\@journalfull{Sexes}\gdef\@journalshort{Sexes}\gdef\@doiabbr{sexes}\gdef\@ISSN{2411-5118}\gdef\@continuouspages{yes}}
\DeclareOption{signals}{\gdef\@journal{signals}\gdef\@journalfull{Signals}\gdef\@journalshort{Signals}\gdef\@doiabbr{signals}\gdef\@ISSN{2624-6120}\gdef\@continuouspages{yes}}
\DeclareOption{sinusitis}{\gdef\@journal{sinusitis}\gdef\@journalfull{Sinusitis}\gdef\@journalshort{Sinusitis}\gdef\@doiabbr{sinusitis}\gdef\@ISSN{2309-107X}\gdef\@continuouspages{yes}}
\DeclareOption{skins}{\gdef\@journal{skins}\gdef\@journalfull{Skins}\gdef\@journalshort{Skins}\gdef\@doiabbr{skins}\gdef\@ISSN{}}
\DeclareOption{smartcities}{\gdef\@journal{smartcities}\gdef\@journalfull{Smart Cities}\gdef\@journalshort{Smart Cities}\gdef\@doiabbr{smartcities}\gdef\@ISSN{2624-6511}\gdef\@continuouspages{yes}}
\DeclareOption{sna}{\gdef\@journal{sna}\gdef\@journalfull{Sinusitis and Asthma}\gdef\@journalshort{Sinusitis Asthma}\gdef\@doiabbr{sna}\gdef\@ISSN{2624-7003}}
\DeclareOption{societies}{\gdef\@journal{societies}\gdef\@journalfull{Societies}\gdef\@journalshort{Societies}\gdef\@doiabbr{soc}\gdef\@ISSN{2075-4698}}
\DeclareOption{socsci}{\gdef\@journal{socsci}\gdef\@journalfull{Social Sciences}\gdef\@journalshort{Soc. Sci.}\gdef\@doiabbr{socsci}\gdef\@ISSN{2076-0760}}
\DeclareOption{software}{\gdef\@journal{software}\gdef\@journalfull{Software}\gdef\@journalshort{Software}\gdef\@doiabbr{software}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{soilsystems}{\gdef\@journal{soilsystems}\gdef\@journalfull{Soil Systems}\gdef\@journalshort{Soil Syst.}\gdef\@doiabbr{soilsystems}\gdef\@ISSN{2571-8789}}
\DeclareOption{solar}{\gdef\@journal{solar}\gdef\@journalfull{Solar}\gdef\@journalshort{Solar}\gdef\@doiabbr{solar}\gdef\@ISSN{2673-9941}\gdef\@continuouspages{yes}}
\DeclareOption{solids}{\gdef\@journal{solids}\gdef\@journalfull{Solids}\gdef\@journalshort{Solids}\gdef\@doiabbr{solids}\gdef\@ISSN{2673-6497}\gdef\@continuouspages{yes}}
\DeclareOption{spectroscj}{\gdef\@journal{spectroscj}\gdef\@journalfull{Spectroscopy Journal}\gdef\@journalshort{Spectrosc. J.}\gdef\@doiabbr{spectroscj}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{sports}{\gdef\@journal{sports}\gdef\@journalfull{Sports}\gdef\@journalshort{Sports}\gdef\@doiabbr{sports}\gdef\@ISSN{2075-4663}}
\DeclareOption{standards}{\gdef\@journal{standards}\gdef\@journalfull{Standards}\gdef\@journalshort{Standards}\gdef\@doiabbr{standards}\gdef\@ISSN{2305-6703}\gdef\@continuouspages{yes}}
\DeclareOption{stats}{\gdef\@journal{stats}\gdef\@journalfull{Stats}\gdef\@journalshort{Stats}\gdef\@doiabbr{stats}\gdef\@ISSN{2571-905X}\gdef\@continuouspages{yes}}
\DeclareOption{std}{\gdef\@journal{std}\gdef\@journalfull{Surgical Techniques Development}\gdef\@journalshort{Surg. Tech. Dev.}\gdef\@doiabbr{std}\gdef\@ISSN{2038-9582}\gdef\@continuouspages{yes}}
\DeclareOption{stresses}{\gdef\@journal{stresses}\gdef\@journalfull{Stresses}\gdef\@journalshort{Stresses}\gdef\@doiabbr{stresses}\gdef\@ISSN{2673-7140}\gdef\@continuouspages{yes}}
\DeclareOption{surfaces}{\gdef\@journal{surfaces}\gdef\@journalfull{Surfaces}\gdef\@journalshort{Surfaces}\gdef\@doiabbr{surfaces}\gdef\@ISSN{2571-9637}\gdef\@continuouspages{yes}}
\DeclareOption{surgeries}{\gdef\@journal{surgeries}\gdef\@journalfull{Surgeries}\gdef\@journalshort{Surgeries}\gdef\@doiabbr{surgeries}\gdef\@ISSN{2017-2017}\gdef\@continuouspages{yes}}
\DeclareOption{suschem}{\gdef\@journal{suschem}\gdef\@journalfull{Sustainable Chemistry}\gdef\@journalshort{Sustain. Chem.}\gdef\@doiabbr{suschem}\gdef\@ISSN{2673-4079}\gdef\@continuouspages{yes}}
\DeclareOption{sustainability}{\gdef\@journal{sustainability}\gdef\@journalfull{Sustainability}\gdef\@journalshort{Sustainability}\gdef\@doiabbr{su}\gdef\@ISSN{2071-1050}}
\DeclareOption{symmetry}{\gdef\@journal{symmetry}\gdef\@journalfull{Symmetry}\gdef\@journalshort{Symmetry}\gdef\@doiabbr{sym}\gdef\@ISSN{2073-8994}}
\DeclareOption{synbio}{\gdef\@journal{synbio}\gdef\@journalfull{SynBio}\gdef\@journalshort{SynBio}\gdef\@doiabbr{synbio}\gdef\@ISSN{2674-0583}\gdef\@continuouspages{yes}}
\DeclareOption{systems}{\gdef\@journal{systems}\gdef\@journalfull{Systems}\gdef\@journalshort{Systems}\gdef\@doiabbr{systems}\gdef\@ISSN{2079-8954}}
\DeclareOption{targets}{\gdef\@journal{targets}\gdef\@journalfull{Targets}\gdef\@journalshort{Targets}\gdef\@doiabbr{targets}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{taxonomy}{\gdef\@journal{taxonomy}\gdef\@journalfull{Taxonomy}\gdef\@journalshort{Taxonomy}\gdef\@doiabbr{taxonomy}\gdef\@ISSN{2673-6500}\gdef\@continuouspages{yes}}
\DeclareOption{technologies}{\gdef\@journal{technologies}\gdef\@journalfull{Technologies}\gdef\@journalshort{Technologies}\gdef\@doiabbr{technologies}\gdef\@ISSN{2227-7080}}
\DeclareOption{telecom}{\gdef\@journal{telecom}\gdef\@journalfull{Telecom}\gdef\@journalshort{Telecom}\gdef\@doiabbr{telecom}\gdef\@ISSN{2673-4001}\gdef\@continuouspages{yes}}
\DeclareOption{test}{\gdef\@journal{test}\gdef\@journalfull{Test}\gdef\@journalshort{Test}\gdef\@doiabbr{}\gdef\@ISSN{}}
\DeclareOption{textiles}{\gdef\@journal{textiles}\gdef\@journalfull{Textiles}\gdef\@journalshort{Textiles}\gdef\@doiabbr{textiles}\gdef\@ISSN{2673-7248}\gdef\@continuouspages{yes}}
\DeclareOption{thalassrep}{\gdef\@journal{thalassrep}\gdef\@journalfull{Thalassemia Reports}\gdef\@journalshort{Thalass. Rep.}\gdef\@doiabbr{thalassrep}\gdef\@ISSN{2039-4365}\gdef\@continuouspages{yes}}
\DeclareOption{thermo}{\gdef\@journal{thermo}\gdef\@journalfull{Thermo}\gdef\@journalshort{Thermo}\gdef\@doiabbr{thermo}\gdef\@ISSN{2673-7264}\gdef\@continuouspages{yes}}
\DeclareOption{timespace}{\gdef\@journal{timespace}\gdef\@journalfull{Time and Space}\gdef\@journalshort{Time Space}\gdef\@doiabbr{timespace}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{tomography}{\gdef\@journal{tomography}\gdef\@journalfull{Tomography}\gdef\@journalshort{Tomography}\gdef\@doiabbr{tomography}\gdef\@ISSN{2379-139X}\gdef\@continuouspages{yes}}
\DeclareOption{tourismhosp}{\gdef\@journal{tourismhosp}\gdef\@journalfull{Tourism and Hospitality}\gdef\@journalshort{Tour. Hosp.}\gdef\@doiabbr{tourhosp}\gdef\@ISSN{2673-5768}\gdef\@continuouspages{yes}}
\DeclareOption{toxics}{\gdef\@journal{toxics}\gdef\@journalfull{Toxics}\gdef\@journalshort{Toxics}\gdef\@doiabbr{toxics}\gdef\@ISSN{2305-6304}}
\DeclareOption{toxins}{\gdef\@journal{toxins}\gdef\@journalfull{Toxins}\gdef\@journalshort{Toxins}\gdef\@doiabbr{toxins}\gdef\@ISSN{2072-6651}}
\DeclareOption{transplantology}{\gdef\@journal{transplantology}\gdef\@journalfull{Transplantology}\gdef\@journalshort{Transplantology}\gdef\@doiabbr{transplantology}\gdef\@ISSN{2673-3943}\gdef\@continuouspages{yes}}
\DeclareOption{transportation}{\gdef\@journal{transportation}\gdef\@journalfull{Transportation}\gdef\@journalshort{Transportation}\gdef\@doiabbr{transportation}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{traumacare}{\gdef\@journal{traumacare}\gdef\@journalfull{Trauma Care}\gdef\@journalshort{Trauma Care}\gdef\@doiabbr{traumacare}\gdef\@ISSN{2673-866X}\gdef\@continuouspages{yes}}
\DeclareOption{traumas}{\gdef\@journal{traumas}\gdef\@journalfull{Traumas}\gdef\@journalshort{Traumas}\gdef\@doiabbr{traumas}\gdef\@ISSN{2673-7477}\gdef\@continuouspages{yes}}
\DeclareOption{tropicalmed}{\gdef\@journal{tropicalmed}\gdef\@journalfull{Tropical Medicine and Infectious Disease}\gdef\@journalshort{Trop. Med. Infect. Dis.}\gdef\@doiabbr{tropicalmed}\gdef\@ISSN{2414-6366}}
\DeclareOption{universe}{\gdef\@journal{universe}\gdef\@journalfull{Universe}\gdef\@journalshort{Universe}\gdef\@doiabbr{universe}\gdef\@ISSN{2218-1997}}
\DeclareOption{urbansci}{\gdef\@journal{urbansci}\gdef\@journalfull{Urban Science}\gdef\@journalshort{Urban Sci.}\gdef\@doiabbr{urbansci}\gdef\@ISSN{2413-8851}}
\DeclareOption{uro}{\gdef\@journal{uro}\gdef\@journalfull{Uro}\gdef\@journalshort{Uro}\gdef\@doiabbr{uro}\gdef\@ISSN{2673-4397}\gdef\@continuouspages{yes}}
\DeclareOption{vaccines}{\gdef\@journal{vaccines}\gdef\@journalfull{Vaccines}\gdef\@journalshort{Vaccines}\gdef\@doiabbr{vaccines}\gdef\@ISSN{2076-393X}}
\DeclareOption{vehicles}{\gdef\@journal{vehicles}\gdef\@journalfull{Vehicles}\gdef\@journalshort{Vehicles}\gdef\@doiabbr{vehicles}\gdef\@ISSN{2624-8921}\gdef\@continuouspages{yes}}
\DeclareOption{venereology}{\gdef\@journal{venereology}\gdef\@journalfull{Venereology}\gdef\@journalshort{Venereology}\gdef\@doiabbr{venereology}\gdef\@ISSN{2674-0710}\gdef\@continuouspages{yes}}
\DeclareOption{vetsci}{\gdef\@journal{vetsci}\gdef\@journalfull{Veterinary Sciences}\gdef\@journalshort{Vet. Sci.}\gdef\@doiabbr{vetsci}\gdef\@ISSN{2306-7381}}
\DeclareOption{vibration}{\gdef\@journal{vibration}\gdef\@journalfull{Vibration}\gdef\@journalshort{Vibration}\gdef\@doiabbr{vibration}\gdef\@ISSN{2571-631X}\gdef\@continuouspages{yes}}
\DeclareOption{virtualworlds}{\gdef\@journal{virtualworlds}\gdef\@journalfull{Virtual Worlds}\gdef\@journalshort{Virtual Worlds}\gdef\@doiabbr{virtualworlds}\gdef\@ISSN{}\gdef\@continuouspages{yes}}
\DeclareOption{viruses}{\gdef\@journal{viruses}\gdef\@journalfull{Viruses}\gdef\@journalshort{Viruses}\gdef\@doiabbr{v}\gdef\@ISSN{1999-4915}}
\DeclareOption{vision}{\gdef\@journal{vision}\gdef\@journalfull{Vision}\gdef\@journalshort{Vision}\gdef\@doiabbr{vision}\gdef\@ISSN{2411-5150}}
\DeclareOption{waste}{\gdef\@journal{waste}\gdef\@journalfull{Waste}\gdef\@journalshort{Waste}\gdef\@doiabbr{waste}\gdef\@ISSN{2813-0391}\gdef\@continuouspages{yes}}
\DeclareOption{water}{\gdef\@journal{water}\gdef\@journalfull{Water}\gdef\@journalshort{Water}\gdef\@doiabbr{w}\gdef\@ISSN{2073-4441}}
\DeclareOption{wem}{\gdef\@journal{wem}\gdef\@journalfull{Wildlife Ecology and Management}\gdef\@journalshort{Wildl. Ecol. Manag.}\gdef\@doiabbr{}\gdef\@ISSN{1234-4321}}
\DeclareOption{wevj}{\gdef\@journal{wevj}\gdef\@journalfull{World Electric Vehicle Journal}\gdef\@journalshort{World Electr. Veh. J.}\gdef\@doiabbr{wevj}\gdef\@ISSN{2032-6653}}
\DeclareOption{wind}{\gdef\@journal{wind}\gdef\@journalfull{Wind}\gdef\@journalshort{Wind}\gdef\@doiabbr{wind}\gdef\@ISSN{2674-032X}\gdef\@continuouspages{yes}}
\DeclareOption{women}{\gdef\@journal{women}\gdef\@journalfull{Women}\gdef\@journalshort{Women}\gdef\@doiabbr{women}\gdef\@ISSN{2673-4184}\gdef\@continuouspages{yes}}
\DeclareOption{world}{\gdef\@journal{world}\gdef\@journalfull{World}\gdef\@journalshort{World}\gdef\@doiabbr{world}\gdef\@ISSN{2673-4060}\gdef\@continuouspages{yes}}
\DeclareOption{youth}{\gdef\@journal{youth}\gdef\@journalfull{Youth}\gdef\@journalshort{Youth}\gdef\@doiabbr{youth}\gdef\@ISSN{2673-995X}\gdef\@continuouspages{yes}}
\DeclareOption{zoonoticdis}{\gdef\@journal{zoonoticdis}\gdef\@journalfull{Zoonotic Diseases}\gdef\@journalshort{Zoonotic Dis.}\gdef\@doiabbr{zoonoticdis}\gdef\@ISSN{2813-0227}\gdef\@continuouspages{yes}}

### 2023-10-31 Proceedings/2023-QIP/QuantumMeasureSubmission.tex
Adding mdpi template and formatted version

--- 2023-10-28 Proceedings/2023-QIP/QuantumMeasure.tex
@@
-\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}
-%\documentclass[aps,pra,10pt,twocolumn,floatfix,nofootinbib]{revtex4-1}
-%\documentclass[10pt,twocolumn,letterpaper]{article}
+%  LaTeX support: latex@mdpi.com 
+%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
 
+%=================================================================
+\documentclass[entropy,article,submit,pdftex,moreauthors]{Definitions/mdpi}
+
+%--------------------
+% Class Options:
+%--------------------
+%----------
+% journal
+%----------
+% Choose between the following MDPI journals:
+% acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arm, arthropoda, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, cyber, dairy, data, ddc, dentistry, dermato, dermatopathology, designs, devices, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, future, futureinternet, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, grasses, gucdd, hazardousmatters, healthcare, hearts, hemato, hematolrep, heritage, higheredu, highthroughput, histories, horticulturae, hospitals, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijpb, ijtm, ijtpp, ime, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jcto, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd, jzbg, kidneydial, kinasesphosphatases, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing,\gdef\@continuouspages{yes}} nanomaterials, ncrna, ndt, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, %%nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, %oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology,\gdef\@ISSN{2813-0618}\gdef\@continuous pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, platforms, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, receptors, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, sclerosis, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, spectroscj, sports, standards, stats, std, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, targets, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis
+% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal. Changing "submit" to "accept" before posting will remove line numbers.
+
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by: 
+% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
+% supfile = supplementary materials
+
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.
+
+%=================================================================
+% MDPI internal commands - do not modify
+\firstpage{1} 
+\makeatletter 
+\setcounter{page}{\@firstpage} 
+\makeatother
+\pubvolume{1}
+\issuenum{1}
+\articlenumber{0}
+\pubyear{2023}
+\copyrightyear{2023}
+%\externaleditor{Academic Editor: Firstname Lastname}
+\datereceived{ } 
+\daterevised{ } % Comment out if no revised date
+\dateaccepted{ } 
+\datepublished{ } 
+%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
+%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
+\hreflink{https://doi.org/} % If needed use \linebreak
+%\doinum{}
+%\pdfoutput=1 % Uncommented for upload to arXiv.org
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
+% cleveref: load \crefname definitions after \begin{document}
 \usepackage{assumptionsofphysics}
@@
 \usepackage{breakurl}
-
-%\usepackage{amsmath}
-%\usepackage{mathrsfs}
-%\usepackage{amsfonts}
-%
-%\usepackage{graphicx}
-%\usepackage{hyperref}
-%\hypersetup{
-%	colorlinks=true,
-%	citecolor=blue,
-%	urlcolor=blue,
-%	linkcolor=blue
-%}
-%\urlstyle{same}
-%\frenchspacing
-%
 \def\>{\rangle}
@@
 
-\begin{document}
+%=================================================================
+% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
 
-\title{How quantum mechanics requires non-additive measures}
-\author{Gabriele Carcassi}
-\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
-\author{Christine A. Aidala}
-\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{How quantum mechanics requires non-additive measures}
 
-\date{\today}
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{Title}
 
-\begin{abstract}
-	Measure theory is used in physics, not just to capture classical probability, but also to quantify the number of states.
+% Author Orchid ID: enter ID or remove command
+\newcommand{\orcidauthorA}{0000-0002-1071-6251} % Add \orcidA{} behind the author's name
+\newcommand{\orcidauthorB}{0000-0001-9540-4988} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Carcassi Gabriele $^{1,*}$\orcidA{} and Christine A. Aidala $^{1}$\orcidB{}}
+
+%\longauthorlist{yes}
+
+% MDPI internal command: Authors, for metadata in PDF
+\AuthorNames{Carcassi Gabriele and Christine A. Aidala}
+
+% MDPI internal command: Authors, for citation in the left column
+\AuthorCitation{Carcassi, C.; Aidala, C. A.}
+% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address[1]{%
+$^{1}$ \quad Physics Department, University of Michigan, Ann Arbor, MI 48109}
+
+% Contact information of the corresponding author
+\corres{Correspondence: carcassi@umich.edu}
+
+% Current address and/or shared authorship
+%\firstnote{Current address: Affiliation 3.} 
+%\secondnote{These authors contributed equally to this work.}
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+%\simplesumm{} % Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\) 
+\abstract{Measure theory is used in physics, not just to capture classical probability, but also to quantify the number of states.
 In previous works, we found that state quantification plays a foundational role in classical mechanics, and therefore, we set ourselves to construct the quantum equivalent of the Liouville measure.
@@
 In this article we show these preliminary results and outline a new line of inquiry that may provide a different insight into the foundations of quantum theory.
-Additionally, this new approach may prove to be useful to those interested in a quantized theory of space-time, as we believe this requires a quantized measure for the quantification of the independent degrees of freedom.
-\end{abstract}
+Additionally, this new approach may prove to be useful to those interested in a quantized theory of space-time, as we believe this requires a quantized measure for the quantification of the independent degrees of freedom.}
 
-\maketitle
+% Keywords
+\keyword{Quantum mechanics; measure theory; non-additive measures; information theory; statistical mechanics}
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Diversity
+%\LSID{\url{http://}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences
+%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data
+%\dataset{DOI number or link to the deposited data set if the data set is published separately.
+If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors.
+In this case, please submit the data set as a supplement.}
+%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Toxins
+%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Encyclopedia
+%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Advances in Respiratory Medicine
+%\addhighlights{yes}
+%\renewcommand{\addhighlights}{%
+
+%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars.
+Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it.
+Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
+%\textbf{What are the main findings?}
+% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
+% \item First bullet.
+% \item Second bullet.
+% \end{itemize}\vspace{3pt}
+%\textbf{What is the implication of the main finding?}
+% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
+% \item First bullet.
+% \item Second bullet.
+% \end{itemize}
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{document}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\section*{Acknowledgments}
-This paper is part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
-This article was made possible through the support of grant \#62847 from the John Templeton Foundation.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\vspace{6pt}
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}
+
+% Only for journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
+A supporting video article is available at doi: link.}
+
+% Only for journal Hardware:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.\vspace{6pt}\\
+%\begin{tabularx}{\textwidth}{lll}
+%\toprule
+%\textbf{Name} & \textbf{Type} & \textbf{Description} \\
+%\midrule
+%S1 & Python script (.py) & Script of python source code used in XX \\
+%S2 & Text (.txt) & Script of modelling code used to make Figure X \\
+%S3 & Text (.txt) & Raw data from experiment X \\
+%S4 & Video (.mp4) & Video demonstrating the hardware in use \\
+%... & ... & ... \\
+%\bottomrule
+%\end{tabularx}
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{Both authors contributed to all stages of the work. All authors have
+	read and agreed to the published version of the manuscript.}
+
+\funding{This article was made possible through the support of grant \#62847 from the John Templeton Foundation.}
+
+\institutionalreview{Not applicable.}
+
+\informedconsent{Not applicable.}
+
+\dataavailability{This study generated no data.}
+
+% Only for journal Nursing Reports
+%\publicinvolvement{Please describe how the public (patients, consumers, carers) were involved in the research.
+Consider reporting against the GRIPP2 (Guidance for Reporting Involvement of Patients and the Public) checklist.
+If the public were not involved in any aspect of the research add: ``No public involvement in any aspect of this research''.}
+
+% Only for journal Nursing Reports
+%\guidelinesstandards{Please add a statement indicating which reporting guideline was used when drafting the report.
+For example, ``This manuscript was drafted against the XXX (the full name of reporting guidelines and citation) for XXX (type of research) research''.
+A complete list of reporting guidelines can be accessed via the equator network: \url{https://www.equator-network.org/}.}
+
+\acknowledgments{This paper is part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.}
+
+\conflictsofinterest{The authors declare no conflict of interest.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{adjustwidth}{-\extralength}{0cm}
+%\printendnotes[custom] % Un-comment to print a list of endnotes
+
+\reftitle{References}
+
+% Please provide either the correct journal abbreviation (e.g.
+according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
+% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.
+
+%=====================================
+% References, variant A: external bibliography
+%=====================================
 \bibliography{bibliography}
 
-\newcommand{\pj}[1] {\underbar{$#1$}}
+% If authors have biography, please use the format below
+%\section*{Short Biography of Authors}
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
+%{\textbf{Firstname Lastname} Biography of first author}
+%
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
+%{\textbf{Firstname Lastname} Biography of second author}
 
+% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
+% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
+% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p.
+328; \citeyear{ref-journal-3b}, p.475).
+This produces: Wong (1999, p.
+328; 2000, p.
+475)
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\PublishersNote{}
+\end{adjustwidth}
 \end{document}

### 2023-10-31 Proceedings/2023-QIP/template.tex
Adding mdpi template and formatted version

--- 2023-10-31 Proceedings/2023-QIP/QuantumMeasureSubmission.tex
@@
 % cleveref: load \crefname definitions after \begin{document}
-\usepackage{assumptionsofphysics}
-\usepackage{tikz}
-\usepackage{breakurl}
-\def\>{\rangle}
-\def\<{\langle}
-\DeclareMathOperator{\erf}{erf}
 
@@
 % Full title of the paper (Capitalized)
-\Title{How quantum mechanics requires non-additive measures}
+\Title{Title}
 
@@
 % Author Orchid ID: enter ID or remove command
-\newcommand{\orcidauthorA}{0000-0002-1071-6251} % Add \orcidA{} behind the author's name
-\newcommand{\orcidauthorB}{0000-0001-9540-4988} % Add \orcidB{} behind the author's name
+\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
+%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
 
 % Authors, for the paper (add full first names)
-\Author{Carcassi Gabriele $^{1,*}$\orcidA{} and Christine A. Aidala $^{1}$\orcidB{}}
+\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}
 
@@
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Carcassi Gabriele and Christine A. Aidala}
+\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}
 
 % MDPI internal command: Authors, for citation in the left column
-\AuthorCitation{Carcassi, C.; Aidala, C. A.}
+\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
@@
 % Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address[1]{%
-$^{1}$ \quad Physics Department, University of Michigan, Ann Arbor, MI 48109}
+\address{%
+$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
+$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}
 
 % Contact information of the corresponding author
-\corres{Correspondence: carcassi@umich.edu}
+\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}
 
 % Current address and/or shared authorship
-%\firstnote{Current address: Affiliation 3.} 
-%\secondnote{These authors contributed equally to this work.}
+\firstnote{Current address: Affiliation 3.} 
+\secondnote{These authors contributed equally to this work.}
 % The commands \thirdnote{} till \eighthnote{} are available for further notes
@@
 % Abstract (Do not insert blank lines, i.e. \\) 
-\abstract{Measure theory is used in physics, not just to capture classical probability, but also to quantify the number of states.
-In previous works, we found that state quantification plays a foundational role in classical mechanics, and therefore, we set ourselves to construct the quantum equivalent of the Liouville measure.
-Unlike the classical counterpart, this quantized measure is non-additive and has a unitary lower bound (i.e.
-no set of states can have less than one state).
-Conversely, requiring that state quantification is finite for finite continuous regions and that each state counts as one already implies non-additivity, which in turn implies the failure of classical theory.
-In this article we show these preliminary results and outline a new line of inquiry that may provide a different insight into the foundations of quantum theory.
-Additionally, this new approach may prove to be useful to those interested in a quantized theory of space-time, as we believe this requires a quantized measure for the quantification of the independent degrees of freedom.}
+\abstract{A single paragraph of about 200 words maximum.
+For research articles, abstracts should give a pertinent overview of the work.
+We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations.
+The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}
 
 % Keywords
-\keyword{Quantum mechanics; measure theory; non-additive measures; information theory; statistical mechanics}
+\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\setcounter{section}{-1} %% Remove this when starting to work on the template.
+\section{How to Use this Template}
+
+The template details the sections that can be used in a manuscript.
+Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section).
+Please check the instructions on the authors' page of the journal to verify the correct order and names.
+For any questions, please contact the editorial office of the journal or support@mdpi.com.
+For LaTeX-related questions please contact latex@mdpi.com.%\endnote{This is an endnote.} % To use endnotes, please un-comment \printendnotes below (before References).
+Only journal Laws uses \footnote.
+
+% The order of the section titles is different for some journals. Please refer to the "Instructions for Authors” on the journal homepage.
 
@@
 
-It is well-established that standard measure theoretic tools cannot be used to characterize quantum probability as quantum theory is a non-classical (i.e.
-non-Kolmogorov) probability theory.
-To obviate this problem, various alternatives and extensions to measure theory and probability theory have been developed with different goals.\cite{groenewold1946principles, moyal1949quantum, gleason1957measures, sorkin1994quantum, hamhalter2003quantum, gudder2009quantum, svozil2022extending, monchietti2023measure} Our interest in measure theory, however, is not in its use to capture classical probability.
-Measures play another crucial use in physical theories: ``counting'', or better ``quantifying'', states.
-It is this use that will be the focus of this work.
+The introduction should briefly place the study in a broad context and highlight why it is important.
+It should define the purpose of the work and its significance.
+The current state of the research field should be reviewed carefully and key publications cited.
+Please highlight controversial and diverging hypotheses when necessary.
+Finally, briefly mention the main aim of the work and highlight the principal conclusions.
+As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research.
+Citing a journal paper \cite{ref-journal}.
+Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}.
+Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Materials and Methods}
 
-As part of our overall project Assumptions of Physics,\cite{aop-book} we have developed an approach, called reverse physics,\cite{aop-phys-ReversePhysics} which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive them.
-We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.\cite{aop-phil-Hamiltonianinformation,aop-phil-HamiltonianPrivilege} The former (i.e.
-the symplectic structure) is the only structure that allows one to quantify states and configurations over independent degrees of freedom in a way that is frame independent.
-If one imposes deterministic and reversible evolution, state quantification must be preserved, leading to Hamiltonian evolution (i.e.
-a symplectomorphism).
-All elements of classical mechanics, including canonical variables, the Lagrangian and the action principle, \cite{aop-phys-ActionPrincipleInterpretation} can be rederived and understood in terms of quantifying the flow of states.
-Given the formal similarity between classical and quantum mechanics, we believe a similar approach is possible for the latter.
-This raises the question: what is the quantum analogue of the Liouville measure, the measure that quantifies states in the classical theory?
+Materials and Methods should be described with sufficient details to allow others to replicate and build on published results.
+Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers.
+Please disclose at the submission stage any restrictions on the availability of materials or information.
+New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.
 
-Statistical mechanics and information theory mandate a link between entropy and state quantification, as the Shannon/Gibbs entropy over uniform distributions corresponds to the logarithm of the count of states.
-In the quantum case, we can start with the von Neumann entropy and define a measure that has the same connection.
-What we find is that this measure is non-additive and that its features can help us better understand what quantization is, why it requires both contextuality and the failure of classical probability, and why the latter can be recovered separately within each context.
-Therefore we believe studying this object is useful not only for the above stated goals of reverse physics, but more in general to get a better understanding of the foundations of quantum mechanics.
+Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers.
+If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review.
+They must be provided prior to publication.
 
-The aim of this article is to propose and articulate the goals of this new line of research.
-It provides the overall picture in which future works will add further mathematical detail and physical insights.
-In section II we will show how measure theory is used to quantify elements in a set and how this is needed to assign probability and probability densities to each element.
-In section III we will see how statistical mechanics and information theory mandate a link between entropy and state quantification, which we will use to define the quantum analogue of the Liouville measure.
-We will find that this measure is not additive.\footnote{Technically, measures are defined to be additive, therefore what we find is not a measure, but a set function.
-However, this term is not widespread, so we use the oxymoron ``non-additive measure'' as it is customary in the mathematical literature.
-\cite{denneberg1994non,nonadditive2014}} In section IV we will see that this non-additivity is linked to non-contextuality.
-In section V we will see that quantization means putting a unitary lower bound on the quantification of states, while keeping a finite value over finite regions.
-We will see that this very directly requires non-additivity and therefore non-contextuality.
-In section VI we will argue that quantization of space-time means using a quantized measure on the quantification of degrees of freedom.
-Lastly, in section VII we will outline a series of goals for a research program that aims to fully characterize the quantized measure.
+Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
+\begin{quote}
+This is an example of a quote.
+\end{quote}
 
-\section{Measure theory and quantifying states}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results}
 
-As we said in the introduction, measure theory is not only used to characterize Kolmogorovian probability.
-More generally, measures are used to define the size of sets.
-Even in probability theory, since probability measures assign probability to sets of points, in order to associate probability to individual points one needs two measures: the \textbf{probability measure} $p : \Sigma_X \to \mathbb{R}$ and another measure quantifying the extent of the set, which we call \textbf{quantifying measure} $\mu : \Sigma_X \to \mathbb{R}$.\footnote{Measures assign values to elements of a $\sigma$-algebra, which can be thought of as a set of experimental procedures.
-\cite{aop-book}} Since we are interested in physical theories, the measure $\mu$ quantifies the number of states.
-Given a set $U$, the ratio $\frac{p(U)}{\mu(U)}$ gives us the average probability per unit of states.
-As we shrink $U$ to a single point $x$, we will find the probability associated to that point, and how this is done depends on the nature of $\mu$.
+This section may be divided by subheadings.
+It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
+\subsection{Subsection}
+\subsubsection{Subsubsection}
 
-In the discrete case, the quantifying measure is simply the counting measure: the cardinality of the set.
-Therefore for a single point we simply have $\frac{p(\{x\})}{\mu(\{x\})} = \frac{p(\{x\})}{1} = p(x)$, which corresponds to the probability associated to the state $x$.
-The counting measure is fully determined by two properties: countable additivity, a defining property of measures, and that each state counts as one.
+Bulleted lists look like this:
+\begin{itemize}
+\item	First bullet;
+\item	Second bullet;
+\item	Third bullet.
+\end{itemize}
 
-In the continuous case, things are a bit different.\footnote{Note that ``discrete' and ``continuous'' have different meanings depending on context.
-At a set theoretic level, ``continuum'' is associated with the cardinality of a set.
-In topology, discrete and continuous spaces are characterized by  properties of open sets, not cardinality: a manifold can be given a discrete topology.
-In the context of quantifying measures in physics, discrete means that we can count elements with natural numbers, while continuous means we use real valued quantities with units, such as meters for distances, to define the size of regions.} If $X$ is a manifold, a set of local coordinates induce a Lebesgue measure, which can be used to quantify the number of possible value combinations.
-Note that a single state has measure zero, unlike the discrete case.
-The Lebesgue measure is defined by two properties: countable additivity and that the size of parallelepipeds corresponds to the volume expressed by the product of the differences of coordinates (i.e.
-$\mu(U) = \Delta x^1 \Delta x^2 \cdots \Delta x^n$).
-Naturally, this is a problem in physics given that the quantification of the number of states should be the same in all frames.
+Numbered lists can be added as follows:
+\begin{enumerate}
+\item	First item; 
+\item	Second item;
+\item	Third item.
+\end{enumerate}
 
-However, the structure of classical phase space (i.e.
-symplectic structure) is exactly the structure that allows one to define a quantifying measure for states that is invariant under change of frame.\cite{aop-book,aop-phil-Hamiltonianinformation,aop-phil-HamiltonianPrivilege} The role of the quantifying measure in classical mechanics, then, is assigned to the Liouville measure, which we can understand as a Lebesgue measure on position and conjugate momentum (i.e.
-canonical coordinates).\footnote{The volume of an infinitesimal parallelepiped is given by $dq^1 dq^2 \cdots dp_1 dp_2 \cdots$.
-Since $dq^i$ is contravariant and $dp_i$ is covariant under change of frame, the volume remains the same.} As we shrink the size of the set, $\frac{p(U)}{\mu(U)}$ becomes the Radon-Nikodym derivative $\frac{dp}{d\mu}$, which gives us the probability density, not the probability itself.
+The text continues here.
 
-%Note that, in both cases, the quantifying measure is potentially unbounded, as the state space can have potentially infinitely many states, as it is the case for one spatial degree of freedom.
-Therefore the quantifying measures are implicitly different from probability measures, and they require specific attention.
+\subsection{Figures, Tables and Schemes}
 
-Quantum mechanics, instead, does not provide a clear state quantifying measure.
-When we characterize preparations and measurement outcomes, we may use the discrete or continuous quantifying measures in different cases for the same system.
-For example, consider a particle.
-If we are interested in a discrete observable, like the energy level of a harmonic oscillator, we would use the counting measure; if we are interested in a continuous observable, like position, we would use the Lebesgue measure associated to the variable.
-As for preparations, the same maximally mixed state for a qubit can be written as a discrete mixture $\frac{1}{2} | 0 \> \< 0 | + \frac{1}{2} | 1 \> \< 1 | $ or as a continuous uniform mixture over all possible states $\int \frac{1}{4\pi} | \psi \> \< \psi | d\Omega$.
+All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, etc.
 
-To be precise, the quantifying measures we considered are really on preparations and measurements, not the states themselves.
-It was the discrete or continuous character of the space of preparations that selected the appropriate space with the appropriate measure.
-But even if we have different preparation and measurement spaces, the state space of the system should always be the same, with a single way to count states.
-This, in other words, highlights the problem of quantum probabilities: while the spaces of preparations and measurement outcomes are both classical, they cannot be connected by classical means.
-That is, we cannot simply put a classical probability of transition between the two because we would be connecting qualitatively different quantifying measures.
+\begin{figure}[H]
+\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
+\caption{This is a figure.
+Schemes follow the same formatting.
+If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel.
+(\textbf{b}) Description of what is contained in the second panel.
+Figures should be placed in the main text near to the first time they are cited.
+A caption on a single line should be centered.\label{fig1}}
+\end{figure}   
+\unskip
 
-To sum up, the typical foundational question is: how do we generalize probability measures for quantum mechanics?
-We are asking something different: given that state-quantifying measures play an equally fundamental role, how do we quantify the number of states in quantum mechanics?
+\begin{table}[H] 
+\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCC}
+\toprule
+\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+\midrule
+Entry 1		& Data			& Data\\
+Entry 2		& Data			& Data \textsuperscript{1}\\
+\bottomrule
+\end{tabularx}
+\noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
+\end{table}
 
-\section{Measure theory and quantifying states}
+The text continues here (Figure~\ref{fig2} and Table~\ref{tab2}).
 
-While the standard structure of quantum mechanics does not give us an answer to our question, statistical mechanics and quantum information theory have an implicit answer.
-Statistical mechanics tells us that the thermodynamic entropy of a system, which is a physically meaningful quantity that has numerous empirical consequences, corresponds to the Shannon entropy for a classical discrete system, the Gibbs entropy\footnote{By Gibbs entropy we mean the differential Shannon entropy $-\int \rho \log \rho d\mu$ calculated over phase space using position and momentum (i.e.
-canonical coordinates).} for a classical continuous system, and the von Neumann entropy for a quantum system.
+% Example of a figure that spans the whole page width. The same concept works for tables, too.
+\begin{figure}[H]
+\begin{adjustwidth}{-\extralength}{0cm}
+\centering
+\includegraphics[width=15.5cm]{Definitions/logo-mdpi}
+\end{adjustwidth}
+\caption{This is a wide figure.\label{fig2}}
+\end{figure}
 
-There is another expression for entropy, often referred to as the fundamental postulate of statistical mechanics, \cite{Peliti2011} which is the logarithm of the count of microstates: $\log \mu(U)$.\footnote{Note that the Boltzmann constant is left out, as entropy will always be expressed in bits or nats.
-Physically, the Boltzmann constant is needed to define temperature in Kelvin.
-Statistical mechanics can be equivalently formulated in terms of $\beta = \frac{1}{k_B T}$.\cite{chyla2011evolution}} In classical statistical mechanics, the measure is indeed the quantifying measure discussed above: the counting measure for the discrete case and the Liouville volume in the continuous case.
-The two expressions for entropy are linked as the Shannon/Gibbs entropy over a uniform distribution matches the logarithm of the measure of the support.
-That is, if $\rho_U$ is a uniform distribution over $U$, then $S(\rho_U) = \log \mu(U)$.
-In quantum mechanics, as we said before, we do not have a measure to quantify states.
-There is, however, a somewhat related expression: the maximum entropy attainable by a quantum system corresponding to a Hilbert space $\mathcal{H}$ is the logarithm of the dimensionality $\dim_{\mathbb{C}}(\mathcal{H})$ of the space.
-Therefore, $ S(\rho_U) \leq \log (\dim_{\mathbb{C}}(\text{span}(U)))$.
+\begin{table}[H]
+\caption{This is a wide table.\label{tab2}}
+	\begin{adjustwidth}{-\extralength}{0cm}
+		\newcolumntype{C}{>{\centering\arraybackslash}X}
+		\begin{tabularx}{\fulllength}{CCCC}
+			\toprule
+			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
+			\midrule
+\multirow[m]{3}{*}{Entry 1 *}	& Data			& Data			& Data\\
+			  	                   & Data			& Data			& Data\\
+			             	      & Data			& Data			& Data\\
+                   \midrule
+\multirow[m]{3}{*}{Entry 2}    & Data			& Data			& Data\\
+			  	                  & Data			& Data			& Data\\
+			             	     & Data			& Data			& Data\\
+                   \midrule
+\multirow[m]{3}{*}{Entry 3}    & Data			& Data			& Data\\
+			  	                 & Data			& Data			& Data\\
+			             	    & Data			& Data			& Data\\
+                  \midrule
+\multirow[m]{3}{*}{Entry 4}   & Data			& Data			& Data\\
+			  	                 & Data			& Data			& Data\\
+			             	    & Data			& Data			& Data\\
+			\bottomrule
+		\end{tabularx}
+	\end{adjustwidth}
+	\noindent{\footnotesize{* Tables may have a footer.}}
+\end{table}
 
-A similar connection can be established in information theory.
-The Shannon/Gibbs/von Neumann entropy quantifies the minimum number of bits required on average to send a message from a given source.\footnote{For a more general characterization of entropy that works in statistical mechanics, information theory and other contexts, and a more precise discussion of the continuous case, see \cite{aop-phys-variability}.} This, again, is a quantity that leads to empirical consequences.
-If we have $n$ bits at our disposal, we have $2^n$ possible combinations, possible messages, that can be encoded.
-Therefore the logarithm of the number of potential messages corresponds to the number of bits that can encode those messages.
-Physically, we can associate a different state to each message to be encoded.
-Therefore $\log \mu(U)$ corresponds to the number of bits.
-This will exactly correspond to the number of bits required by a source characterized by a uniform probability distribution over that number of symbols.
+%\begin{listing}[H]
+%\caption{Title of the listing}
+%\rule{\columnwidth}{1pt}
+%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
+%\rule{\columnwidth}{1pt}
+%\end{listing}
 
-While the above discussion does not cover all technical details, the only point we want to make is that the link $S(\rho_U) = \log \mu(U)$ is fully justified by both statistical mechanics and information theory.
-It is not a capricious request, but rather a necessary requirement stemming not from just one, but from two disciplines that have empirical consequences.
-Therefore if quantum mechanics does not directly provide us with a measure to quantify states, we can use what it does provide, the von Neumann entropy, to define and calculate said measure.
+Text.
 
-We can start by calculating the simplest case, which is when $U = \{ \psi \}$ is a single state.
-In this case, we have $S(|\psi\>\<\psi|) = 0$ and therefore $\mu(\{\psi\}) = 2^{0} = 1$.
-The measure returns one for each state as in the classical discrete case.
+Text.
 
-The next case is when $U = \{ \psi, \phi \}$ is the set of two states. A uniform distribution over two states will correspond to the mixed state
+\subsection{Formatting of Mathematical Components}
+
+This is the example 1 of equation:
+\begin{linenomath}
 \begin{equation}
-	\rho = \frac{1}{2} |\psi\>\<\psi| + \frac{1}{2} |\phi\>\<\phi|.
+a = 1,
 \end{equation}
-The entropy of said state will depend on the probability $p=|\<\psi | \phi \>|^2$ in the following way
+\end{linenomath}
+the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
+%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
+This ensures correct linenumbering.
+The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
+
+This is the example 2 of equation:
+\begin{adjustwidth}{-\extralength}{0cm}
 \begin{equation}
-	S(\rho) = - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} 
-	- \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2}.
+a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
 \end{equation}
-If the states are orthogonal, then $p=0$, and $S(\rho) = 1$ therefore $\mu(U) = 2^1 = 2$.
-In this case we have that $2 = \mu(\{ \psi, \phi \}) = \mu(\{\psi\}) + \mu(\{\phi\}) = 1 + 1$, and therefore $\mu$ is additive.
-Moreover, if $U$ is an orthonormal basis, the von Neumann entropy $S(U) = \log(\dim_{\mathbb{C}}(\text{span}(U))) = \log |U|$ is the logarithm of the cardinality of $U$.
-Therefore the quantifying measure $\mu(U) = |U|$ is equal to the counting measure.
-But this is a special case that corresponds to the highest entropy reachable with a mixture.
-In all other cases the entropy will be lower, and therefore, in general, $\mu(\{ \psi, \phi \}) \leq \mu(\{\psi\}) + \mu(\{\phi\})$.
-In other words, the measure $\mu$ is not additive.
+\end{adjustwidth}
 
-To sum up, the physically motivated way to quantify the number of states in quantum mechanics is, in general, not additive. When counting states in quantum mechanics, $1+1 \leq 2$.
+% Example of a page in landscape format (with table and table footnote).
+%\startlandscape
+%\begin{table}[H] %% Table in wide page
+%\caption{This is a very wide table.\label{tab3}}
+%	\begin{tabularx}{\textwidth}{CCCC}
+%		\toprule
+%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
+%		\midrule
+%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
+%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
+%		\bottomrule
+%	\end{tabularx}
+%	\begin{adjustwidth}{+\extralength}{0cm}
+%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
+%	\end{adjustwidth}
+%\end{table}
+%\finishlandscape
 
-\section{Non-additivity and contextuality}
+Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
+%% Example of a theorem:
+\begin{Theorem}
+Example text of a theorem.
+\end{Theorem}
 
-While unusual claims are common in quantum mechanics, we want to have a clear, physically tenable, reason as to why the quantifying measure for quantum states must be non-additive.
-In the context of a spin 1/2 system, for example, why is it that $\{ |z^+\>, |z^-\>\}$ count as two states while $\{ |z^+\>, |x^+\> \}$ count as less than two?
+The text continues here. Proofs must be formatted as follows:
 
-Since we saw that the quantifying measure is indeed additive for a set of orthogonal states, this should make us think about contextuality: the impossibility to assign outcomes to all potential measurements of a quantum system.
-We have, in fact, derived that it is impossible to define an additive quantifying measure on all potential states, but we can do so on the outcomes of one measurement.
-The ability/inability to define additive measures is exactly the difference between classical/non-classical probability.
+%% Example of a proof:
+\begin{proof}[Proof of Theorem 1]
+Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
+\end{proof}
+The text continues here.
 
-While contextuality is often presented as a somewhat abstract, almost metaphysical, property of quantum mechanics, we have a rather more down-to-earth view.
-The choice of the measurement means a difference in physical process of either the preparation or measurement device.
-Choosing which direction of spin to prepare or measure means, at some point, orienting some polarizer, magnetic field or making an equivalent change to a physical device.
-Something is changed in the physical world, not just in the system, but around the system, in its environment.
-If Amanda claims that a particle is in the $|z^+\>$ state, then Boris is able to infer something about Amanda's preparation/measuring device.\footnote{This is not at all particular to quantum mechanics.
-We can study the motion of a cannonball here on earth, but not on the surface of the sun, as the cannonball would be immediately vaporized.
-The mere claim that a system exists implies that the boundary conditions necessary for that system to exist are satisfied.
-In this sense, all physics is contextual.} Therefore we can understand the context simply as those conditions on the environment that are necessary for the system to be found in that particular state.
-A set of orthogonal states, then, has the property that they can be found in the same external conditions.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Discussion}
 
-Given that we have, in general, a correlation between system and environment, it is natural to understand how contextuality leads to a non-additive measure.
-Two states, in fact, should be counted as two only if they lead to a change in the system alone.
-That is, \textbf{states should be quantified all-else-being-equal}.
-While two states from different contexts are indeed two distinct states, stating we transitioned from one to the other does not give us information only about the system: it tells us something about the environment as well.
-The degree of incompatibility of observables, then, ultimately measures the degree of change when realizing the context physically.
+Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses.
+The findings and their implications should be discussed in the broadest context possible.
+Future research directions may also be highlighted.
 
-So we have a clear, physically motivated reason as to why contextuality is linked to a non-additive measure.
-States should be quantified at-all-else-being-equal, meaning that possible correlations with the environment have to be removed.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusions}
 
-\section{Non-additivity and quantization}
+This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.
 
-While we have clarified what it means for the quantifying measure to be non-additive, and how this is tied to non-contextuality, we haven't shown why it should be linked to quantization.
-In fact, what exactly is quantization?
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Patents}
 
-Before trying to answer that question, let's make a direct comparison between the quantifying measures for the three cases: the counting measure $\mu_d$ for the classical discrete case, the Liouville measure $\mu_c$ for the classical continuous case and the newly defined quantized measure $\mu_q$ for the quantum case.
-This requires a space where all three can be defined, which fortunately exists: the surface of a sphere.
-First, we want to compare the behavior on a single point.
-The counting measure returns one, and so does the quantum measure.
-The Liouville measure, however, returns zero.
-Second, we want to compare the behavior on a finite region, an open set.
-The Liouville measure returns the solid angle, which is always going to be finite as it must be less than or equal to $4\pi$.
-The von Neumann entropy of a mixed state over a two-state system is between zero and one, and therefore the measure must be between one and two.
-However, given that an open set on a sphere has infinitely many points, the counting measure over a finite region will be infinite.
-Lastly, we want to compare the additivity.
-Both classical measures are additive, while the quantum measure is not.
-
-Let us list the three properties we have mentioned:
-\begin{enumerate}
-	\item single states count as one
-	\item finite continuous regions correspond to finite state quantification
-	\item state quantification is additive on disjoint regions.
-\end{enumerate}
-These seem very reasonable properties to assume for a quantifying measure of states.
-\textbf{However, these three conditions are incompatible}.
-If each state counts as one, and the measure is additive, then a finite continuous region must be infinite as it contains infinite points.
-\textbf{Of the three conditions, we can only pick two}.
-This is, in fact, what each measure does.
-The counting measure rejects the second property.
-The Liouville (and Lebesgue) measure discards the first.
-The quantum measure sacrifices the third.
-
-So, in light of this, what is quantization?
-If we have a system characterized by continuous quantities, such as position, the state space will necessarily be a continuous space and finite regions will have to correspond to finite information as those correspond to the finite precision measurements we can make.\footnote{For a treatment of how finite-time experimental verification is captured by topological concepts, and what necessary and sufficient operational assumptions are required for the use of real valued quantities, see \cite{kelly1996,aop-book}.} If the regions are large enough, the Liouville (or Lebesgue) measure will work.
-But as we shrink a region we will encounter, given the additivity, regions that contain less than one state, which does not make sense.
-Note that measure less than one corresponds to negative thermodynamic entropy, which does not make sense either.
-
-Quantization, then, is fixing the quantifying measure so that it is bounded from below by one (and the entropy by zero): each state counts as one.
-However, this comes at a price: non-additivity.
-This non-additivity is ultimately why classical probability has to fail as well, given that additivity is one of its axioms.
-It is also the same reason as why quantization implies contextuality: the non-additivity means that some states cannot be prepared at-all-else-being-equal.
-
-In our view, this provides a good conceptual basis to understand the failure of classical mechanics, the need for the departure from classical ideas and the need for quantum theory.
-We believe this, by itself, is already sufficient reason to start a research program to explore these ideas fully.
-However, since most physicists do not seem to be interested in this type of ``clean-up work'', we can easily argue that a full characterization of the quantized measure is critical for the development of future physical theories.
-
-\section{Implication for space-time quantization}
-
-We have seen that quantization can be understood as putting a unitary lower bound on the state quantification: no set of states can have fewer than one state.
-If we were to quantize space-time, what is it that we are quantizing?
-
-The most fundamental theories that are available to us are field theories.
-In a field theory, the state is described by the value of the fields at each point in space.
-At each point, the value of the field is independent from the others: the degrees of freedom form a continuum.\footnote{If we restrict ourselves to continuous fields, then we only need to specify the value of the fields on a countable dense subset (i.e.
-rational values of the coordinates).
-However, these details do not change the discussion: a Lebesgue measure on the rationals has the same characteristics as the one extended to the reals.} We can therefore pose the question: how do we quantify the degrees of freedom?
-
-If we have a region of space and we double the volume, we can imagine we are also doubling the number of degrees of freedom.
-Therefore, we can construct a volume measure from the metric tensor, which would give us an invariant volume that quantifies degrees of freedom, much in the same way that the Liouville measure quantified states.
-
-Since the measure is finite for finite regions and is additive, as we shrink the region, we will find regions with less than one degree of freedom.
-This does not make sense.
-As before, we need to put a unitary lower bound on the measure and our measure cannot be additive.
-We need a quantized measure.
-Quantizing space-time, then, means putting a quantized measure on the number of degrees of freedom.
-
-What does it mean that two degrees of freedom count less than two?
-Before answering that question, let's consider time.
-We said that if we double the spatial volume, we double the number of degrees of freedom.
-Suppose we have a space-time region and we double the time, are we doubling the number of degrees of freedom?
-Well, we are certainly doubling the support of the function we have to specify for the field configuration, but we wouldn't consider that doubling the degrees of freedom.
-Intuitively, we just think of those as being the same degrees of freedom that are getting different values in time.
-This is because we expect the future and past values to be linked by some dynamical equations, so the future values would be linked, in some way, to the past values.
-That is, the past and future value of the same field at some point are not independent.
-This may give us an insight into the physical meaning of the non-additivity of quantification of the number of degrees of freedom.
-
-The same field at two spatial points may count less than two degrees of freedom simply because their values are not independent.
-The values of the field at two distant points can be chosen arbitrarily, therefore the measure would be additive at long range.
-But as the points get sufficiently close, this would not be the case, so at short scale we need a quantized measure to quantify independent degrees of freedom.
-Note that the lack of independence of the field values in close-by regions may mean that high frequencies of the fields are damped: the values have to become closer.
-A quantized measure of the degrees of freedom would effectively provide a cutoff.
-Therefore we have a story that is very well motivated and provides a possible solution for high energy divergences in quantum field theories.
-
-While the development of a quantum theory of gravity is not our primary interest, the mathematical tools required to characterize the non-additive quantized measure seem to be of similar nature as the ones that may be used to quantize space-time.
-
-\section{Open questions and future work}
-
-Having outlined the scope and motivation for this new approach, we now outline the main questions we think need to be addressed by this new research program.
-
-Both the counting measure and the Lebesgue/Liouville measures are unique in some sense.
-For the counting measure, once we define the value for a single point, typically one, the whole measure is determined.
-Similarly, once we have defined a real variable, the Lebesgue measure is determined.
-The Liouville measure requires a bit more structure, but once the canonical coordinates are chosen, the measure is determined as well.
-Therefore the first question is whether these quantized measures are also unique in some sense.
-
-While we have found some literature on non-additive measures and set functions \cite{hamhalter2003quantum, nonadditive2014}, together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increases.
-Unfortunately, the quantized measure does not have this feature.
-Consider the set $\{ |z^+\>, |z^-\> \}$.
-An equal mixture will return the maximally mixed state which, as we said, would correspond to a measure equal to 2.
-Now consider the set $\{ |z^+\>, |z^-\>, |x^+\> \}$.
-An equal mixture of them will not correspond to the maximally mixed state, and therefore the measure will be less than two.
-The set is bigger, but the measure is smaller.
-Therefore there may be new mathematical ground to be explored.
-
-We may want to understand the necessary conditions for a quantifying measure.
-Some may be derived from the strict concavity of entropy, given the connection between the two.
-Furthermore, note that the pairwise relationships between the measure over two points, the entropy of the equal mixture of two states and the probability of transition between two states are all invertible.
-This means that the relationship between two states defines the geometry of the whole state space.
-Therefore there must be further consistency requirements that can be identified.
-
-Another issue that needs to be understood is how the quantized measure behaves in composite systems and across multiple degrees of freedom.
-In the classical case, the measure is defined on the product of the corresponding $\sigma$-algebras.
-Ideally, we would like to show that this simple procedure cannot support a quantized measure, and the $\sigma$-algebra of the tensor product must be taken.\footnote{Technically it would be the $\sigma$-algebra projective space coming from the tensor product of the Hilbert spaces associated with the individual systems.}
-
-The first spaces we should study are the symplectic ones as these are the ones that allow for frame independent densities.
-If we want to extend to a quantization of space-time, however, Riemannian spaces need to be studied as well.
-However, we have found a connection between the two: if the symplectic form is expressed in position and velocity, instead of canonical coordinates, its position-velocity component corresponds to the metric tensor.
-The metric tensor can then be understood as defining volumes in position-velocity instead in position-momentum.
-This may allow us to transfer the quantized measure onto the tangent bundle and then, hopefully, to the space itself.
-
-Finally, we want to stress that, contrary to what happens in other programs, the physical motivations for the mathematical generalization are tightly driven by the physics.
-We are not introducing concepts that, like negative probability, have no clear direct physical meaning.
-The goal is the quantification of states.
-
-\section{Conclusion}
-
-We have seen that, in physics, measure theory is not only used to characterize classical probability but also to quantify states in state spaces.
-Ideally, such a measure would have the following three conditions:
-\begin{enumerate}
-	\item single states count as one
-	\item finite continuous regions correspond to finite state quantification
-	\item state quantification is additive on disjoint regions.
-\end{enumerate}
-However, these conditions are not compatible, and only two of the three can be chosen.
-The classical discrete case discards the second; the classical continuous case discards the first; the quantum case discards the third.
-Quantization, then, can be understood as putting a unitary lower bound on the measure that quantifies states.
-
-The lack of additivity on the whole space makes it impossible to use classical probability on the state space as a whole.
-Physically, the issue is that states must be counted at-all-else-being-equal, a feature broken by contextuality.
-But if the states do belong to the same context (e.g.
-they are the output of the same measurement), additivity is recovered.
-This tells us very directly why quantization requires contextuality, and why classical probability works within the same context.
-The fact that single states and finite continuous regions both have finite measure also tells us why quantum mechanics mixes features of the classical discrete and continuous cases.
-
-We believe that further study and characterization of the non-additive quantized measure will not only bring more clarity to the interplay among the above concepts, but it may be necessary for the development of future physical theories.
-Given that, in a field theory, the number of independent degrees of freedom is taken to be proportional to the spatial volume, quantization of space-time can be understood as the quantization of the measure that quantifies the independent degrees of freedom.
-The development of physically motivated mathematical tools may help in that regard.
+This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{Both authors contributed to all stages of the work. All authors have
-	read and agreed to the published version of the manuscript.}
+\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided.
+The following statements should be used ``Conceptualization, X.X.
+and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y.
+and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y.
+All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation.
+Authorship must be limited to those who have contributed substantially to the work~reported.}
 
-\funding{This article was made possible through the support of grant \#62847 from the John Templeton Foundation.}
+\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''.
+Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}
 
-\institutionalreview{Not applicable.}
+\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study.
+You might choose to exclude this statement if the study did not require ethical approval.
+Please note that the Editorial Office might ask you for further information.
+Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans.
+OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals.
+OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}
 
-\informedconsent{Not applicable.}
+\informedconsent{Any research article describing a study involving humans should contain this statement.
+Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans.
+You might also choose to exclude this statement if the study did not involve humans.
 
-\dataavailability{This study generated no data.}
+Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
+Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}
+
+\dataavailability{We encourage all authors of articles published in MDPI journals to share their research data.
+In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study.
+Where no new data were created, or where data is unavailable due to privacy or ethical restrictions, a statement is still required.
+Suggested Data Availability Statements are available in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.}
 
@@
 
-\acknowledgments{This paper is part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.}
+\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections.
+This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}
 
-\conflictsofinterest{The authors declare no conflict of interest.}
+\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results.
+Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section.
+If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results''.}
 
@@
 %% Optional
+
+%% Only for journal Encyclopedia
+%\entrylink{The Link to this entry published on the encyclopedia platform.}
+
+\abbreviations{Abbreviations}{
+The following abbreviations are used in this manuscript:\\
+
+\noindent 
+\begin{tabular}{@{}ll}
+MDPI & Multidisciplinary Digital Publishing Institute\\
+DOAJ & Directory of open access journals\\
+TLA & Three letter acronym\\
+LD & Linear dichroism
+\end{tabular}
+}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
+\appendix
+\section[\appendixname~\thesection]{}
+\subsection[\appendixname~\thesubsection]{}
+The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+
+\begin{table}[H] 
+\caption{This is a table caption.\label{tab5}}
+\newcolumntype{C}{>{\centering\arraybackslash}X}
+\begin{tabularx}{\textwidth}{CCC}
+\toprule
+\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
+\midrule
+Entry 1		& Data			& Data\\
+Entry 2		& Data			& Data\\
+\bottomrule
+\end{tabularx}
+\end{table}
+
+\section[\appendixname~\thesection]{}
+All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.
 
@@
 %=====================================
-\bibliography{bibliography}
+%\bibliography{your_external_BibTeX_file}
+
+%=====================================
+% References, variant B: internal bibliography
+%=====================================
+\begin{thebibliography}{999}
+% Reference 1
+\bibitem[Author1(year)]{ref-journal}
+Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
+% Reference 2
+\bibitem[Author2(year)]{ref-book1}
+Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
+% Reference 3
+\bibitem[Author3(year)]{ref-book2}
+Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
+% Reference 4
+\bibitem[Author4(year)]{ref-unpublish}
+Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
+% Reference 5
+\bibitem[Author5(year)]{ref-communication}
+Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
+% Reference 6
+\bibitem[Author6(year)]{ref-proceeding}
+Author 1, A.B.; Author 2, C.D.; Author 3, E.F.
+Title of presentation.
+In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
+% Reference 7
+\bibitem[Author7(year)]{ref-thesis}
+Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
+% Reference 8
+\bibitem[Author8(year)]{ref-url}
+Title of Site. Available online: URL (accessed on Day Month Year).
+\end{thebibliography}
 

### 2023-11-21 Articles/2023-HilbertQuantum/PhysScripta/response1.tex
Adding appeal draft reviewed by C and G

\documentclass[11pt, executivepaper]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{enumitem}
\usepackage{geometry}
 \geometry{
 a4paper,
 total={158mm,237mm},
 left=28mm,
 top=28mm,
 }
\usepackage{hyperref}
\hypersetup{colorlinks= true, allcolors=blue}
\setcitestyle{aysep={}}
\begin{document}

\title{\textbf{Replies to Reviewers' Comments}}

\author{The Authors}

\maketitle

\noindent First of all, we would like to thank the editorial board for the consideration of our article. Stepping back, the main line of inquiry of this paper, and our work more in general, is determining whether a particular mathematical structure is the ``correct'' one to represent a physical system. While we are convinced that no real progress can be made in the foundations of physics without establishing a precise link between mathematical models and the physical systems they are supposed to represent, the academic community does not seem set up to encourage these discussions. Since the problem sits right in between mathematics and physics, most physicists do not have enough of the technical mathematical background required, and most mathematicians do not have the physical intuition that in the end must guide the decisions.

For example, one of us is working on the ill-definition of the path integral formulation of quantum field theory. That is, one of the most successfully used frameworks in physics assumes a measure that is mathematically proven not to exist. While one may think this is an important problem to address, most mathematicians consider the discussion settled that there is no solution, without looking for ways to address the problem, and most physicists simply do not care given that the calculations can be made to work, sometimes boasting of the fact that they do not care.

As another example, two of us have worked on the physical significance of the Lagrangian and the action principle, clarifying what the action is, and why, under specific physical assumptions, it is minimized. The paper was rejected without review by a good number of journals, including one specialized in the foundations of physics, claiming that the topic is not of general interest. It was published in a ``catch-all'' journal. Given this context, we are grateful that the current paper was at least sent for review.

In the cover letter we specifically recommended ``not relying exclusively on reviewers who would focus primarily on the mathematical aspects'' because ``the core new idea is to look for physical justification of the mathematical structure, not simply internal mathematical consistency.'' We found, in fact, that the vast majority of physicists who read the paper concurred that having a time evolution oscillating between finite and infinite expectation values is an objective problem. On the other hand, most mathematicians either abstained (as it is a physical matter), or argued that the mathematics is consistent, and therefore there was no problem. Both reviewers, in different manners, did that, and as predicted, ``missed the main point of the paper.''

We are now going to look at the reviews, section by section, to show how that is the case.

\subsection*{Reviewer 1}

Reviewer 1 starts with
\begin{quote}
This article is partially in my areas of expertise. The main observations I
would like to make are the following.
Some called paradoxes in quantum mechanics have been solved with proper
use of mathematical formalism of non-relativistic quantum mechanics. See,
in particular, the introduction of the book: D.M. Gitman, I.V. Tyutin, and
B.L. Voronov. Self-adjoint Extensions in Quantum Mechanics: General Theory
and Applications to Schrodinger and Dirac Equations with Singular Potentials,
volume 62 of Progress in Mathematical Physics. Birkhauser, Boston, 2012. I
think that references of this kind are related to the scope and subjects of the
article submitted by the authors.
\end{quote}

The introduction of Gitman-Tyutin-Voronov clearly spells out what type of problems it aims to address:
\begin{quote}
The problem of a correct definition of quantum observables is generally nontrivial in the case of physical systems with boundaries and/or with singular interactions (including QFT models). In what
follows, for the sake of brevity, we call such systems nontrivial physical systems (or simply nontrivial systems). The interest in this problem revives periodically in connection with studies of specific nontrivial systems such as a particle on a
finite interval or on a semiaxis, a particle in singular potential fields, in particular in the Aharonov–Bohm or in $\delta$-like potential fields, and so on.
\end{quote}
That is, if one goes past the ``trivial'' case in which the domain of integrable functions is the entire $\mathbb{R}^n$, different issues arise, such as the difference between Hermitian, self-adjoint and symmetric operators. Note that the problem of infinite/undefined expectations exists also in the trivial case. Contrary to what the reviewer says, these references do not address the issue at hand any more than the references we already provided.

The reviewer continues
\begin{quote}
	On the other hand, in Section C, the authors provide two examples of wave
	functions. From the elementary definition of an operator, we need to specify its
	domain and the rule that assigns the image of each element of its domain. In
	particular, if we had the operator $x^n$ in $L^2$, the maximal domain of definition
	is $\{\psi \in L^2 : x^n \psi \in L^2\}$. The wave function given in Equation 1 belongs to any
	of these domains, for n = 0, 1, . . . ; however, the wave function in Equation 2
	will not belong to such domains since $n \geq 1$. Henceforth, the statement that
	``expectation values will diverge for all moments'' does not make sense because
	the wave function given in Equation 1 does not belong to the domain of any
	power $x^n$ for $n \geq 1$.
\end{quote}
The reviewer points out that, mathematically, it would be more proper to say that the expectation of the observable is not defined, instead of being infinite. Therefore, we should say that the expectation value of the energy, or the position, or the number of particles, is not defined. How does this solve the physical problem? Arguably, saying the expectation of the energy of a particular state is undefined is even worse than saying that it is infinite.

\begin{quote}
	Another subject that I want to touch on is unitary transformations in nonrelativistic
	quantum mechanics.

	Recall that an operator from one Hilbert space onto another is said to be
	unitary or isomorphic if defined on the whole of the first one, sends it onto the
	second Hilbert space, and preserves the inner product. See, for example, Section
	40, p. 112, of N.I. Akhiezer and I.M. Glazman. Theory of Linear Operators in
	Hilbert Space. Dover Publications, New York, 2 edition, 1993.

	The definition of a unitary transformation that I know is used in the mathematical
	formalism of non-relativistic quantum mechanics is the definition above.
	If this definition is equivalent to a change of variable, a reference discussing the
	equivalence of both concepts would benefit the reader.
\end{quote}
We use the same standard definition of unitary transformation. If we didn't, the other reviewer would have raised the same objection. The fact that coordinate transformations, as all other changes of representation, are unitary transformations is established by Wigner's theorem. Even the Wikipedia page on Wigner's theorem mentions passive coordinate transformations in the context of changes of representation:
\begin{quote}
	Loosely speaking, a symmetry transformation is a change in which "nothing happens"[8] or a "change in our point of view"[9] that does not change the outcomes of possible experiments. For example, translating a system in a homogeneous environment should have no qualitative effect on the outcomes of experiments made on the system. Likewise for rotating a system in an isotropic environment. This becomes even clearer when one considers the mathematically equivalent passive transformations, i.e. simply changes of coordinates and let the system be.
\end{quote}

The reviewer continues
\begin{quote}
Along with the unitary transformation in quantum mechanics, we have the
concept of two unitarily equivalent operators: Let H1 and H2 be two Hilbert
spaces. Two linear operators say $\hat{P}$ in H1 and $\hat{S}$ in H2 with domains $D(\hat{P}) \subset H1$
and $D(\hat{S}) \subset H1$, are isomorphic (or unitarily equivalent) if there is an unitary
transformation $F$ such that $D(\hat{S}) = F(D(\hat{P}))$ and $S = F\hat{P}F^{-1}$ hold. See, for
example, the provided definition in p.115 of N.I. Akhiezer and I.M. Glazman.
Theory of Linear Operators in Hilbert Space. Dover Publications, New York, 2
edition, 1993.
With the definition of unitarily equivalent operators via a unitary transformation,
it is prevented from having states with finite expectation values to
be unitarily mapped to states with infinite expectation values, provided that
we start with a state that belongs to the corresponding domain of the power
operator.
\end{quote}
The reviewer confuses old position operator in the new frame $U X U^{-1}$  with the new position operator in the new frame $Y$. The old position operator $X$ and the new position operator $Y$ are \textbf{not} unitarily equivalent under the change of coordinate since the expectation value for $x$ and $y$ for the same state are necessarily going to be different. That's the whole point of a coordinate transformation: the position variable is different in the two frames.

\begin{quote}
The previous issues need to be revisited by the authors since some of their
main conclusions are based on the definitions of unitary transformation and
unitarily equivalent operators, which seem not to be considered by them. If
the authors are dealing with and referring to different concepts, please, provide
the (mathematical) definitions and references where they are introduced and
discussed.
\end{quote}
Note that in the appendix, proposition 24 explicitly constructed the unitary transformation from the change of variables. Also note that the other reviewer didn't encounter a problem following our definitions.

Changes of coordinates are a basic and fundamental notion in physics, and a pre-requisite to follow many physics arguments, including the one we make. The reviewer does not demonstrate a good understanding of the subject. Moreover, the reviewer shows no appreciation for the physical problem: they do not seem concerned that quantities like position, momentum, energy, number of particles, or any other unbounded quantities are not well defined on all states. This is exactly the kind of attitude that, as mentioned above, prevents the advancement of the mathematical foundations of physics.

While we understand and agree that the lack of definition of unbounded quantities over a Hilbert space is not a mathematical problem, the lack of definition of position, momentum, energy or number of particles on a pure state is indeed a physical problem. The review shows that the reviewer is not capable of making that determination.

\subsection*{Reviewer 2}

Reviewer 2 starts with
\begin{quote}
The standard formulation of elementary quantum mechanics is in terms of
self-adjoint operators acting in Hilbert space. The thesis of this paper is that
part of this structure is unphysical.

A Hilbert space is first of all a complex vector space. Second, it has an
inner product. Finally, it is complete as a metric space. The unit vectors in
Hilbert space determine the states of the system. Given a state and a self-adjoint
operator, there is a corresponding probability distribution defined for the Borel
subsets of the real numbers. The most important self-adjoint operators are the
position operator Q and the momentum operator P.

The authors accept the vector space structure and the inner product as
‘physical”. However they reject the completeness as “non-physical”. The authors
justify this on the grounds that “completeness forces us to include states
with an infinite expectation value, which cannot be implemented physically”.
\end{quote}
The reviewer has correctly summed up the main point of the article.  The reviewer then proceeds with the following remark
\begin{quote}
The rather vague notion of “unphysical” used by the authors sometimes has
to do with cardinality and sometime has to do with the extended real number
system. However the main part of the argument involves only the latter notion.
\end{quote}
We defined physicality as
\begin{quote}
	On the other
	hand, a mathematical definition is \textbf{unphysical} if it can
	be shown to require properties or operations that cannot
	have a physical counterpart.
\end{quote}
So, yes, different mathematical requirements may be unphysical depending on the the physics they are trying to describe. Modeling the number of people with positive and negative numbers, for example, is unphysical. Modeling distance with complex numbers is unphysical. In our opinion, the fact that some reviewers of \textbf{physics} papers have a problem with something that most physicists would take as obvious shows the sorry state of mathematical modeling in theoretical physics.

\begin{quote}
The authors give an explicit example. One can take the Hilbert space to
consist of square-integrable functions of x, and Q to be multiplication by x.
Then a state with Gaussian density gives finite expectation to Q, while a state
with Cauchy density has undefined expectation. The Gaussian states gives finite
expectation to Q to an even power, while the state with Cauchy density gives
infinite expectation to Q to an even power.
\end{quote}
The reviewer has summed up the example correctly. This shows that we are not using non-standard techniques, as claimed by the previous reviewer. The reviewer proceeds to argue that having infinite expectation is not a problem.
\begin{quote}
The obvious criticism of this point of view is that exactly the same issue
arises in probability theory. In fact, given a self-adjoint operator A and a state
$\psi$, one may think of A as a random variable with a well-defined distribution.
That is, if S is a Borel subset of the reals, then the indicator function $1_S$ applied
to A is represented as an event $1_S(A)$ with expectation equal to the probability
that A is in S. The physical consequence of this is that in repeated experiments
sample proportions are likely to be close to probabilities.

More generally, if f is a bounded Borel function, then f(A) is realized as a
bounded random variable. The physical consequence is that in repeated trials
the sample means are likely to be close to the expectation. In short, even though
a random variable may have expectation infinity or minus infinity or undefined
(infinity minus infinity), there is no problem extracting useful information from
it.

But what if one says that it is important to consider that the random variable
be directly realized as predicting outcome values in independent repeated trials?
This should be realized by sample means of the random variable itself.

If the expectation is finite, then the sample mean values should be finite and
for a large number of trials be close to the expectation. However even if the
expectation is infinite, the sample means will be finite. All that happens is that
for larger and larger numbers of trials the sample mean values will be larger and
larger. How about the case when the expectation is undefined (infinity minus
infinity)? Again the sample mean values will be finite. But they will fluctuate,
getting arbitrarily positive and arbitrarily negative.

In the case of the Cauchy distribution cited by the authors the sample means
themselves have the Cauchy distribution. There is nothing infinite about that.
\end{quote}
The reviewer is essentially arguing that the infinity comes when you get infinite samples. Every outcome of each trial is finite. This would work if we were considering a statistical mixture of states that have finite expectation. At each trial, we would get a pure state from the mixture that has finite expectation. However, this is not the case.

The state we constructed is a pure state, not a mixed state. That pure state can be itself the outcome of a \textbf{single} trial. The expectation cannot be understood simply as the average of infinitely many trials. If the infinite quantity is energy, given that it must be conserved, the pure state requires infinite energy to be prepared. A superposition of energy eigenstates cannot be understood as a statistical mixture.

Moreover, we have also shown that, in Hilbert spaces, a unitary transformation exists that changes a finite expectation value to an infinite expectation value \textbf{in finite time}. Even if we allow the idea that pure states are themselves ensembles (a matter that is interpretation dependent), you still have this problem. Therefore, at the very least, Hilbert spaces are not physical because they allow such time evolutions.

This provides a good example of the type of confusion which stems from the imprecise mapping between math and physics. Note that this misunderstanding on the part of the reviewer is precisely a problem of inadequate mapping between mathematical objects and physical ones, the type of problem that our research program aims to address.

\begin{quote}
The authors make a positive suggestion: Consider using wave functions in
the Schwartz space of smooth rapidly decreasing functions. For these states all
powers of Q and P have finite expectation.

But it seems awkward to have a restriction that such moments must have
finite expectations, since this rules out most quantum mechanical calculations,
including such central examples as the ground state of hydrogen. In that case
higher moments of momentum diverge.
\end{quote}
The reviewer claims that the restriction to Schwartz spaces is ``awkward'', as it supposedly rules out most quantum mechanical calculations. Here is what we wrote in the conclusion
\begin{quote}
Similarly, we can pose a problem on the
Schwartz space, extend to the Hilbert space or the space
of distributions for calculation, and then bring the result
back to the Schwartz space
\end{quote}
That is, we \textbf{specifically stated} that the space used for calculation can be whatever one wants. The ground state of the hydrogen atom is
$$ \frac{1}{\sqrt{\pi} a_0^{3/2} }e^{-r/a_0} $$
which is a Schwartz function. So the result is a Schwartz function (it decreases to infinity like an exponential).

Note that we gave a \textbf{physical reason} for a possible restriction to the Schwartz space: the ability to perform quantum tomography, to identify states experimentally. Note that the reviewer uses mathematical convenience in calculation to discard this reason. It shows little consideration for the physical argument, which, again, is the main point.

\begin{quote}
The reviewer concludes that the issues discussed in this paper really have to
do with general probability theory, and in this context they seem not to be a
problem.
\end{quote}
Again, there is no issue in probability theory as each element itself has finite expectation. The infinity is potential as it is the expectation over infinitely many trials. Moreover, probability theory is not a theory of states and processes, like quantum mechanics. The issue here is that a single pure state can have an infinite value: the infinity is actual, not potential. Also, time evolutions are included that realize the actual infinity in finite time.

\begin{quote}
There are some questionable statements in the paper. For instance, it is
claimed that the Schwartz space is the only vector subspace closed under Fourier
transformations. This is clearly false: Consider the domain of a power of the
harmonic oscillator Hamiltonian H = (1/2)(P2 + Q2).

There is a claim that a change of variables leads to an antiunitary transformation.
This seems to be a confusion arising from the fact that the authors
neglect the absolute value on the Jacobian determinant associated with the
change of variables.
\end{quote}
These are details that can be addressed, and have no bearing on the argument. Also note that the second criticism is on a detail of the change of variable, which means the reviewer had no problem with the rest, contrary to the first reviewer.

The reviewer concludes
\begin{quote}
A final comment. The reviewer argued above that the way that a random
variable is realized is by doing the experiment (perhaps many times) and examining
the actual outcome values. This is one very standard approach to
interpreting probability.
In quantum mechanics this works for a self-adjoint operator, or even for
a family of commuting self-adjoint operators. However there is the following
result. It is so general that it does not even require an assumption that the
experiment is repeated.
Theorem. Consider a finite-dimensional Hilbert space of dimension at least
4. Suppose that for each self-adjoint operator A there is an “experimental”
value v(A). Suppose these values satisfy at least the following two properties:
• v(A) is always an eigenvalue of A
• A commutes with B implies v(A + B) = v(A) + v(B).
Then this leads to a contradiction. For a proof see Bricmont, Goldstein,
Hemmick: From EPR-Schrodinger paradox to nonlocality based on perfect correlations.
In other words, the restriction to a commuting family is essential. But what
is the precise mathematical criterion for choosing the particular family? That
is the real problem of “physicality”.
\end{quote}
The reviewer now discusses interpretations of probability, and a theorem related to hidden variables. He claims that the real problem of ``physicality'' lies in understanding these issues. This makes it clear that, in the eye of the reviewer, we are not working on ``real'' issues.

Note that hidden variables are, by definition, not experimentally definable. They cannot be measured, prepared, or studied in any way. They are mere conjectures. Therefore, in the context of this paper, they are unphysical. Note that most of the technical appendix of our paper is there to guarantee that all that we say is \textbf{independent} of interpretation. We reformulate superposition in terms of statistical mixtures because the latter have a clear physical meaning. We go out of our way to avoid interpretational issues, so that everything that we say is grounded in experimental constraints.

In our opinion, the reviewer read the paper, understood the main point but, since their personal interests lie elsewhere, promptly dismissed it as irrelevant. Unfortunately, this is the attitude of the majority of people working on the foundations of quantum mechanics: if you are interested in positing hidden variables, parallel worlds, consciousness-induced collapses or other similar ideas, then you find a place in the ongoing discussion; if you want to understand exactly what mathematical features are physically justified, and maybe find mathematical definitions that are better suited to capture experimental/physical requirements, then you are dismissed because you are not working on ``real'' problems.

\subsection*{Conclusion}

The main point of the paper was a physical one: we cannot prepare or measure infinite quantities, therefore pure states identified by infinite quantities cannot be physical. The only mathematical point required is to show that completeness forces us to include pure states that correspond to infinite/undefined quantities. Neither of the two reviewers disputed the mathematical point. Neither of the two reviewers, in our opinion, seriously and knowledgeably considered the physical point.

While the immediate interest is for this paper, we feel that it is more important to recognize and mention the bigger problem, of which these reviews are a particular instance. We strongly believe that working toward mathematical tools that see their foundation/motivation in physical ideas and requirements is the only way to proceed toward a better understanding of our physical theories. This will have practical advantages in better understanding, better intuition and better generalized tools. This cannot happen if we can't even make the absolutely simple point that we cannot have pure states whose position or energy is undefined/infinite.

We are actively looking for a journal in which such line of research can be discussed. We would welcome harsh criticism if it were on point and would push us towards our goal. We realize that it is hard to find reviewers on these topics precisely because, for various reasons, there is little active research on them... but one of the reasons there is little active research on them is because they are very difficult to publish.

Naturally, it is the prerogative of the editorial board to decide whether this line of research is appropriate for the journal, particularly given the additional problem of finding suitable reviewers. We would completely understand if, even if in principle you agree with the spirit of our research, practical considerations make it unrealistic to pursue within the journal. In such case, we only ask you to let us know if you are aware of people, conferences or other journals that would be sympathetic to our endeavor.

%\clearpage
%\bibliographystyle{apalike}
%\bibliography{bibliography}
\end{document}

### 2023-11-25 Articles/2023-HilbertQuantum/PhysScripta/response1.tex
Comments from Francisco

@@
 While we are convinced that no real progress can be made in the foundations of physics without establishing a precise link between mathematical models and the physical systems they are supposed to represent, the academic community does not seem set up to encourage these discussions.
-Since the problem sits right in between mathematics and physics, most physicists do not have enough of the technical mathematical background required, and most mathematicians do not have the physical intuition that in the end must guide the decisions.
+Since the problem sits right in between mathematics and physics, and given that the fields have separate training, conferences, journals, culture, etc.,  most physicists do not have enough of the technical mathematical background required, and most mathematicians do not have the physical intuition that in the end must guide the decisions.
 
@@
 On the other hand, most mathematicians either abstained (as it is a physical matter), or argued that the mathematics is consistent, and therefore there was no problem.
-Both reviewers, in different manners, did that, and as predicted, ``missed the main point of the paper.''
+Both reviewers focused mainly on the math, and as predicted, ``missed the main point of the paper.''
 
@@
 Arguably, saying the expectation of the energy of a particular state is undefined is even worse than saying that it is infinite.
+The carefulness in specifying the domain seems to be motivated by having all the mathematical objects being well-defined, not all the mathematical objects adequately representing the physical entities they intend to represent as is the goal of our paper.
 
@@
 Changes of coordinates are a basic and fundamental notion in physics, and a pre-requisite to follow many physics arguments, including the one we make.
-The reviewer does not demonstrate a good understanding of the subject.
+The reviewer does not demonstrate a good understanding.
 Moreover, the reviewer shows no appreciation for the physical problem: they do not seem concerned that quantities like position, momentum, energy, number of particles, or any other unbounded quantities are not well defined on all states.
@@
 The reviewer now discusses interpretations of probability, and a theorem related to hidden variables.
-He claims that the real problem of ``physicality'' lies in understanding these issues.
+They claim that the real problem of ``physicality'' lies in understanding these issues.
 This makes it clear that, in the eye of the reviewer, we are not working on ``real'' issues.

### 2023-11-25 Articles/2023-HilbertQuantum/Submissions/Physica Scripta/response1.tex
Appeal letter and rebuttal, with rebuttal moved to correct directory
Also original cover letter and initial response to editors, which were not committed previously

--- 2023-11-25 Articles/2023-HilbertQuantum/PhysScripta/response1.tex
@@
 
-\title{\textbf{Replies to Reviewers' Comments}}
+\title{\textbf{Rebuttal to Reviewers' Comments on Manuscript PHYSSCR-125221}}
 
@@
 
-\noindent First of all, we would like to thank the editorial board for the consideration of our article.
-Stepping back, the main line of inquiry of this paper, and our work more in general, is determining whether a particular mathematical structure is the ``correct'' one to represent a physical system.
-While we are convinced that no real progress can be made in the foundations of physics without establishing a precise link between mathematical models and the physical systems they are supposed to represent, the academic community does not seem set up to encourage these discussions.
-Since the problem sits right in between mathematics and physics, and given that the fields have separate training, conferences, journals, culture, etc.,  most physicists do not have enough of the technical mathematical background required, and most mathematicians do not have the physical intuition that in the end must guide the decisions.
-
-For example, one of us is working on the ill-definition of the path integral formulation of quantum field theory.
-That is, one of the most successfully used frameworks in physics assumes a measure that is mathematically proven not to exist.
-While one may think this is an important problem to address, most mathematicians consider the discussion settled that there is no solution, without looking for ways to address the problem, and most physicists simply do not care given that the calculations can be made to work, sometimes boasting of the fact that they do not care.
-
-As another example, two of us have worked on the physical significance of the Lagrangian and the action principle, clarifying what the action is, and why, under specific physical assumptions, it is minimized.
-The paper was rejected without review by a good number of journals, including one specialized in the foundations of physics, claiming that the topic is not of general interest.
-It was published in a ``catch-all'' journal.
-Given this context, we are grateful that the current paper was at least sent for review.
-
-In the cover letter we specifically recommended ``not relying exclusively on reviewers who would focus primarily on the mathematical aspects'' because ``the core new idea is to look for physical justification of the mathematical structure, not simply internal mathematical consistency.'' We found, in fact, that the vast majority of physicists who read the paper concurred that having a time evolution oscillating between finite and infinite expectation values is an objective problem.
-On the other hand, most mathematicians either abstained (as it is a physical matter), or argued that the mathematics is consistent, and therefore there was no problem.
-Both reviewers focused mainly on the math, and as predicted, ``missed the main point of the paper.''
-
-We are now going to look at the reviews, section by section, to show how that is the case.
+The main point of the paper was a physical one: we cannot prepare or measure infinite quantities, therefore pure states identified by infinite quantities cannot be physical.
+The only mathematical point required is to show that completeness forces us to include pure states that correspond to infinite/undefined quantities.
+We believe that the reviews did not appropriately evaluate the manuscript and would like to appeal the decision made by the journal.
+Below we consider the reviews, section by section, to show how both reviewers focused mainly on the math and did not seriously or knowledgeably consider the physics.
 
@@
 
-\subsection*{Conclusion}
-
-The main point of the paper was a physical one: we cannot prepare or measure infinite quantities, therefore pure states identified by infinite quantities cannot be physical.
-The only mathematical point required is to show that completeness forces us to include pure states that correspond to infinite/undefined quantities.
-Neither of the two reviewers disputed the mathematical point.
-Neither of the two reviewers, in our opinion, seriously and knowledgeably considered the physical point.
-
-While the immediate interest is for this paper, we feel that it is more important to recognize and mention the bigger problem, of which these reviews are a particular instance.
-We strongly believe that working toward mathematical tools that see their foundation/motivation in physical ideas and requirements is the only way to proceed toward a better understanding of our physical theories.
-This will have practical advantages in better understanding, better intuition and better generalized tools.
-This cannot happen if we can't even make the absolutely simple point that we cannot have pure states whose position or energy is undefined/infinite.
-
-We are actively looking for a journal in which such line of research can be discussed.
-We would welcome harsh criticism if it were on point and would push us towards our goal.
-We realize that it is hard to find reviewers on these topics precisely because, for various reasons, there is little active research on them...
-but one of the reasons there is little active research on them is because they are very difficult to publish.
-
-Naturally, it is the prerogative of the editorial board to decide whether this line of research is appropriate for the journal, particularly given the additional problem of finding suitable reviewers.
-We would completely understand if, even if in principle you agree with the spirit of our research, practical considerations make it unrealistic to pursue within the journal.
-In such case, we only ask you to let us know if you are aware of people, conferences or other journals that would be sympathetic to our endeavor.
-
 %\clearpage

### 2023-12-10 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
HarriganSpekkens: remove color highlights

--- 2023-10-13 Articles/2021-HarriganSpekkens/HarriganSpekkensTakeTwo.tex
@@
 
-In 2010 Harrigan and Spekkens (HS) proposed a rigorous classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}.
+In 2010 Harrigan and Spekkens (HS) proposed a formal classification in order to categorize the nature of the quantum state, i.e.\ to establish whether in a certain model $\psi$ corresponds to a real property of a quantum object, in which case the model is called $\psi$-ontic, or to some observer information, making it $\psi$-epistemic \cite{Harrigan:2010}.
 While their original aim was to clarify Einstein's view of Quantum Mechanics (QM), the HS framework has been widely employed in the literature not only to categorize different formulations of QM, but also to argue what types of interpretations are admissible (\cite{Leifer:2013, Leifer:2017, Branciard:2014, Hermens:2021, Wood:2015, Ringbauer:2015, Mazurek:2016, Bartlett:2012}; cf.\ \cite{Oldofredi:2020b, Ladyman:2021} for critical discussions).
@@
 
-Such a theorem had a remarkable resonance (\cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}), and questions about its actual meaning are still discussed today: on the one hand, it rules out interpretations of QM where $\psi$ merely represents information.
-On the other, scholars recently showed that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument (\cite{Ben:2017, Rizzi:2018, Oldofredi:2021, DeBrota:2019}).
+Such a theorem had a remarkable resonance \cite{Leifer:2014, Leifer:2014b, Lewis:2012, Renner:2012, Colbeck:2017, Hardy:2013, Maroney:2014, Patra:2013, Mansfield:2016, Schlosshauer:2012, Schlosshauer:2013, Schlosshauer:2014, Aaronson:2013}, and questions about its actual meaning are still discussed today: on the one hand, some authors believe that it rules out interpretations of QM where $\psi$ merely represents information.
+On the other hand, it has recently been shown by other scholars that non-trivial epistemic as well as statistical approaches to QM are not refuted by the PBR argument \cite{Ben:2017, Rizzi:2018, Oldofredi:2020b, Oldofredi:2021, DeBrota:2019}.
 
 While the discussion has been focused mainly on whether $\psi$-epistemic models are problematic or whether the PBR argument itself is problematic, we ask a different question: what if the underlying HS classification itself is problematic?
-In particular, the framework assumes that once pure states are fully characterized, mixed states are constructed simply as a standard (i.e.
-classical) statistical mixture.
+In particular, the framework assumes that once pure states are fully characterized, mixed states are constructed simply as a standard (i.e.\ classical) statistical mixture.
 However, as it is well-known, classical and quantum mixtures have different properties.
@@
 
-Using this new approach, in this paper we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM.
+Employing this new approach, in this paper we show that $\psi$-ontic models cannot reproduce all the predictions, results and implications of QM.
 We start by noting that the von Neumann entropy plays a crucial role in the predictions of both quantum statistical mechanics and quantum information theory.
@@
 
+After having introduced the relevant definitions of the HS ontological model framework, it should be noted that it has been subject to several criticisms, as already said in the previous section.
+For instance, Oldofredi \& Lopez in \cite{Oldofredi:2020b} analyzed the implicit assumptions made by Harrigan \& Spekkens in order to define their categorization, and showed that relevant interpretations of quantum theory---such as the statistical, relational and perspectival readings of the quantum formalism---cannot be classified according to the HS approach, given that their principles are in tension with those employed in \cite{Harrigan:2010}.
+Similarly, Hance, Rarity and Ladyman argued in \cite{Ladyman:2021} that the dichotomy ``ontic vs.\ epistemic'' proposed by HS is artificial, since there are cases in which the quantum state can represent both ontic properties of quantum systems as well as agents' knowledge of them.
+Thus, they claim, the definitions of $\psi$-ontic and $\psi$-epistemic introduced a few lines above do not capture the intuitive ideas with which physicists and philosophers alike use the terms ``ontic'' and ``epistemic''.
+
+While we agree with the main messages of these critical responses to the HS framework, the argument offered in this paper follows a different and complementary strategy.
+In fact, instead of criticizing the assumptions used by Harrigan \& Spekkens, we consider them valid (for the sake of the argument) and derive a no-go theorem for $\psi$-ontic models directly from them, as we are going to show in the next section.
+
 \begin{figure}
@@
 This includes those given by statistical mechanics and information theory, which depend on the von Neumann entropy $H_\mathcal{H} : M(\mathcal{H}) \to \mathbb{R}$.
-Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability.\cite{aop-phys-variability} It is also the one that provides the correct link to the thermodynamic entropy.} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.
+Given that $M(\Lambda)$ is the space of probability measures over $\Lambda$, it is natural to use the Shannon/Gibbs entropy function $H_\Lambda : M(\Lambda) \to \mathbb{R}$, as this is what is typically done in statistical mechanics.\footnote{Note that, while many functions called ``entropy'' exist, the Shannon entropy is the only indicator of variability that is continuous, monotonic and linear in probability \cite{Carcassi:2021}.
+It is also the one that provides the correct link to the thermodynamic entropy.} Since we want to be able to replicate the predictions of statistical mechanics and information theory, a very reasonable request is that these two entropies, calculated on the different representations, agree.
 
@@
 This assumes that a composite system is describable using a standard classical probability measure, which, in turn, allows us to express the joint probability distribution of observables of different subsystems.
-This, in general, will fail because it is ``not quantum enough.''
+Again, this in general will fail because it is ``not quantum enough''.
 
@@
 
-Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new\cite{Werner:2014}.
+Now, the fact that quantum theory does not follow the rules of classical-Kolmogorov probability is nothing new \cite{Werner:2014, Pitowsky:1989}.
 However, this does not seem to be enough to clear confusion on the subject because, admittedly, we do use classical probability in the context of quantum mechanics.
@@
 The temptation, then, is to simply assume that all we need to do is just put the right transition probability between the two spaces, and we are done.
-This is exactly what cannot be done and what doesn't work.
+This is exactly what cannot be done and what does not work.
 That is
@@
 
+\section{On interpretations}
+
+Our analysis entails an interesting consequence for the philosophical discussion concerning the interpretation of quantum theory.
+Indeed, many scholars use the PBR result to vindicate the empirical adequacy of a given framework---especially those in which the quantum state is somehow supposed to represent real physical objects---or to rule out a specific reading of the quantum formalism.
+Here we warn physicists and philosophers that no argument resting on the Harrigan \& Spekkens approach should be employed in such metaphysical debates, given that no theory satisfying the requirements imposed by their categorization is able to reproduce \emph{all} the results of quantum mechanics.
+
+Therefore, we not only agree with Hance, Rarity and Ladyman in saying that the HS definitions of $\psi$-ontic and $\psi$-epistemic models do not capture the actual meaning usually assigned to the terms ``ontic'' and ``epistemic'' as they are employed when interpreting quantum theories, but also say the following:
+\begin{quote}
+	\textbf{If a certain interpretation of the quantum formalism\footnote{As for instance de Broglie-Bohm theory, Everett's relative state formulation, relational quantum mechanics or the many-worlds perspective \emph{etc.}.} is fully empirically equivalent to quantum mechanics, then it cannot be categorized as a $\psi$-ontic model.}
+\end{quote}
+\noindent In turn, the logical conjunction of our result and the PBR theorem entails that no interpretation of quantum theory can be correctly categorized according to the HS approach.
+
+Another philosophical issue affecting the definition of $\psi$-ontic model must be mentioned.
+In their paper Harrigan \& Spekkens require that if a model is $\psi$-ontic, the ontic state $\lambda$ fully identifies the quantum state, i.e.\ only one quantum state is compatible with the ontic state of the system.
+This fact, in turn, means that $\psi$ should be understood realistically, as an entity representing some real physical property of the system, or even the system itself as in $\psi$-complete models.
+However, as interestingly noted by Halvorson, realism towards the quantum state is highly problematic, since each representation relation proposed in the literature entails several conundrums that should be avoided---cf.\  \cite{Halvorson:2019} for details.
+With regard to this, the HS framework does not clearly spell out what is meant by the fact that $\psi$ ``represents'' a real property of physical systems.
+In other words, to claim that in $\psi$-complete models $\lambda$ is uniquely represented by a single quantum state leaves many ontological questions unanswered.
+For instance, it is not specified in which way $\psi$ represents the ontic state: is it a direct or indirect representation?
+If $\psi$ is a real object, as one would expect in $\psi$-complete models, is there a literal identification between $\psi$ and $\lambda$?
+On these and other issues the Harrigan \& Spekkens framework does not give guidance.
+
+Therefore, not only is their categorization shown to be empty, but also it does not offer a rigorous philosophical ground on which to cast the interpretational debate.
+
 \section{Conclusion}
@@
 
-The key problem of the HS categorization is that it is ``not quantum enough.'' The framework tries to hide the non-classicality at the level of each ontic state, which, by construction, cannot be decomposed further using a standard probability measure.
+The key problem of the HS categorization is that it is ``not quantum enough''.
+The framework in fact tries to hide the non-classicality at the level of each ontic state which, by construction, cannot be decomposed further using a standard probability measure.
 However, HS definitions entail that statistical mixtures are simply classical mixtures in the sense that they are captured by a standard probability measure.
-But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.\footnote{The convex space formed by the mixtures is radically different\cite{bengtsson_życzkowski_2017}.
-There is also debate as whether proper and improper mixtures are distinguishable in the context of quantum mechanics\cite{Kirkpatrick:2001,DEspagnat:2001}.} Therefore, the HS categorization prevents one from describing quantum statistical mixing.
+But quantum mixtures are non-classical in the same way that the quantum transition probabilities defined by the inner product are non-classical.\footnote{The convex space formed by the mixtures is radically different \cite{Bengtsson:2017}.
+There is also debate as to whether proper and improper mixtures are distinguishable in the context of quantum mechanics \cite{Kirkpatrick:2001}, \cite{DEspagnat:2001}.} Therefore, the HS categorization prevents one from describing quantum statistical mixing.
 
-Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are untenable not because there is a problem with $\psi$-epistemic models in particular, but rather there is an underlying problem in the HS definitions.
+Our conclusion is that the PBR theorem finds that $\psi$-epistemic models are untenable not because there is a problem with epistemic models in particular, but rather there is an underlying problem in the HS definitions.
 Specifically, the problem of the HS categorization is the use of standard probability theory in composite systems.
 If understood in this manner, the PBR result is the first part of a more general argument showing the inadequacy of the HS framework to categorize quantum theories, an argument that we have completed with the proposed no-go theorem against $\psi$-ontic models.
+The direct consequence of this is that the HS categorization should not be employed as a tool to classify quantum interpretations or to draw philosophically sound conclusions on the nature of quantum mechanics.
 
@@
 derive the Holevo bound).
-Since the latter are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.
+Since these are derived from the more general framework of quantum mechanics, one cannot cherry pick which results a model has to reproduce to be a valid model of quantum theory: \emph{all} results must be correctly reproduced.
 

### 2023-12-15 Proceedings/2023-QIP/QuantumMeasureSubmission.tex
QIP: changes after review

@@
 
-As part of our overall project Assumptions of Physics,\cite{aop-book} we have developed an approach, called reverse physics,\cite{aop-phys-ReversePhysics} which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive them.
-We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.\cite{aop-phil-Hamiltonianinformation,aop-phil-HamiltonianPrivilege} The former (i.e.
+As part of our overall project Assumptions of Physics,\cite{aop-book} we have developed an approach, called reverse physics,\cite{aop-phys-ReversePhysics} which, starting from the laws, aims to find a minimal set of physical assumptions needed to rederive them.\footnote{To see how the reverse physics approach compares and contrasts with respect to other works that aim to find physically motivated starting points, for example in the context of quantum theory,\cite{zeilinger1999,Hardy:2001jk,clifton2003,bub2005,chiribella2011} we refer to \cite{aop-phys-ReversePhysics}.} We found that the ability to quantify states is a crucial component of classical theory, so much so that it is enough to recover the structure of phase space and the laws of evolution.\cite{aop-phil-Hamiltonianinformation,aop-phil-HamiltonianPrivilege} The former (i.e.
 the symplectic structure) is the only structure that allows one to quantify states and configurations over independent degrees of freedom in a way that is frame independent.
@@
 
-Quantum mechanics, instead, does not provide a clear state quantifying measure.
-When we characterize preparations and measurement outcomes, we may use the discrete or continuous quantifying measures in different cases for the same system.
+Quantum mechanics, instead, does not provide a clear state quantifying measure.\footnote{In this context, a pure state, or simply state, is mathematically represented by a ray of a Hilbert space (or equivalently a point in the projective space) and a mixed state is represented by a density matrix (i.e.
+a positive semi-definite Hermitian operator with trace one.} When we characterize preparations and measurement outcomes, we may use the discrete or continuous quantifying measures in different cases for the same system.
 For example, consider a particle.
@@
 
-\section{Measure theory and quantifying states}
+\section{Entropy and state quantification}
 
@@
 While unusual claims are common in quantum mechanics, we want to have a clear, physically tenable, reason as to why the quantifying measure for quantum states must be non-additive.
-In the context of a spin 1/2 system, for example, why is it that $\{ |z^+\>, |z^-\>\}$ count as two states while $\{ |z^+\>, |x^+\> \}$ count as less than two?
+In the context of a spin 1/2 system, for example, why is it that the set of pure states $\{ |z^+\>, |z^-\>\}$ counts as two states while the set of pure states $\{ |z^+\>, |x^+\> \}$ counts as less than two?
 
@@
 This requires a space where all three can be defined, which fortunately exists: the surface of a sphere.
+Note that the comparison is only at the mathematical level, so the only requirement is to find a mathematical space that can be used to represent states in the three cases.
+The sphere is used in classical mechanics to define the orientation of an angular momentum in three dimensions; the manifold is symplectic and allows both a Liouville measure and a counting measure.
+The (Bloch) sphere is used in quantum mechanics to represent the states of a two-state system, which can also represent the orientation of a spin one-half particle, and therefore it also allows a quantized measure.
+
 First, we want to compare the behavior on a single point.
@@
 While we have found some literature on non-additive measures and set functions \cite{hamhalter2003quantum, nonadditive2014}, together with modified theories of integration, these seem to concentrate on monotonic measures: measures that increase as the set increases.
-Unfortunately, the quantized measure does not have this feature.
+Interestingly, the quantized measure does not have this feature.
 Consider the set $\{ |z^+\>, |z^-\> \}$.
@@
 The set is bigger, but the measure is smaller.
-Therefore there may be new mathematical ground to be explored.
+Therefore there may be new mathematical ground to be explored as well as conceptual ground.
+The physical meaning of non-monotonicity is not yet clear to us.
+We know it appears only when adding states that can be expressed as a superposition of the original set.
+One approach\footnote{Also suggested by one anonymous reviewer.} would be to study the set of points that can be added without change to the measure.
 
@@
 This may allow us to transfer the quantized measure onto the tangent bundle and then, hopefully, to the space itself.
+
+Stepping back from the mathematical problem, there are also more conceptual issues that need to be understood.
+For example, what are the operational requirements to be able to quantify the number of states?
+It may turn out that the ability to operationally define and manipulate a system already provides all that is needed to define how to quantify states.
 

### 2024-01-23 Articles/2023-HilbertQuantum/HilbertQuantum.tex
Add \usepackage{breakurl} so bibliography formats correctly on arxiv
HilbertQuantum: removing "only space"

@@
 \usepackage{tikz}
+\usepackage{breakurl}
 
@@
 We suspect that at least some problems in quantum theory related to infinities may be ultimately caused by the wrong space being used.
-We strongly believe a better solution can be found, and we look at Schwartz spaces for inspiration, as, among other things, they guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are the only space closed under Fourier transform.
+We strongly believe a better solution can be found, and we look at Schwartz spaces for inspiration, as, among other things, they guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are closed under Fourier transform.
 \end{abstract}
@@
 We will show that other options exist, offering Schwartz spaces as an example.
-These guarantee the expectation values of all polynomials of position and momentum to be finite, and they are also the only space closed under Fourier transform.
+These guarantee the expectation values of all polynomials of position and momentum to be finite, and they are also closed under Fourier transform.
 
@@
 \item its closure over all Cauchy sequences recovers the Hilbert space - the only thing missing are the Cauchy sequences for which the limit of at least one expectation is not well defined
-\item it is the only space closed under Fourier transform - the Fourier transform of a Schwartz function is a Schwartz function
+\item it is closed under Fourier transform - the Fourier transform of a Schwartz function is a Schwartz function
 \item it plays a fundamental role in the theory of distributions - objects like the $\delta$-function are mathematically defined on top of the Schwartz space
@@
 
-The Schwartz space also has a very important property: it is the only space that is closed under the Fourier transform.
-That is, the Fourier transform of a Schwartz function is a Schwartz function, and it is the only set of functions with this property.
+The Schwartz space also has another important property: it is closed under the Fourier transform.
+That is, the Fourier transform of a Schwartz function is a Schwartz function.
 Physically, this means that we can always express a wavefunction in either the position or the momentum basis.

### 2024-01-23 Articles/2023-HilbertQuantum/HilbertQuantum.tex
HilbertQuantum: removing change of orientation

@@
 \end{equation}
-If the change of variable does not change orientation, the above transformation is unitary.\footnote{If the orientation changes, the transformation is anti-unitary.}
+The above transformation is unitary.
 
@@
 Then the map $U : \mathcal{H} \to \mathcal{H}$ such that
-	$\phi(y) = U(\psi(x)) \mapsto \psi(f^{-1}(y)) \sqrt{\left(\frac{df}{dx}\right)^{-1}}$ is a unitary transformation.
+	$\phi(y) = U(\psi(x)) \mapsto \psi(f^{-1}(y)) \sqrt{\left|\frac{df}{dx}\right|^{-1}}$ is a unitary transformation.
 \end{prop}
@@
 			\int \phi^\dagger(y) \phi(y) dy &= \int U(\psi(x))^\dagger U(\psi(x)) dy \\
-			 = \int \psi^\dagger(f^{-1}(y)) &\sqrt{\left(\frac{df}{dx}\right)^{-1}}\psi(f^{-1}(y)) \sqrt{\left(\frac{df}{dx}\right)^{-1}} dy \\
-			 &= \int \psi^\dagger(f^{-1}(y)) \psi(f^{-1}(y)) \left(\frac{df}{dx}\right)^{-1} dy \\
+			 = \int \psi^\dagger(f^{-1}(y)) &\sqrt{\left|\frac{df}{dx}\right|^{-1}}\psi(f^{-1}(y)) \sqrt{\left|\frac{df}{dx}\right|^{-1}} dy \\
+			 &= \int \psi^\dagger(f^{-1}(y)) \psi(f^{-1}(y)) \left|\frac{df}{dx}\right|^{-1} dy \\
 			 &= \int \psi^\dagger(x(y)) \psi(x(y)) \frac{dx}{dy} dy \\

### 2024-05-18 Proceedings/2024-IPMU/carcassi-quantummeasure.tex
Add IPMU paper

%%%%%%%%%%%%%%%%%%%% author.tex %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% sample root file for your "contribution" to a proceedings volume
%
% Use this file as a template for your own input.
%
%%%%%%%%%%%%%%%% Springer %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass{styles/svproc}
%
% RECOMMENDED %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\usepackage{amsfonts}
\def\UrlFont{\rmfamily}
\def\>{\rangle}
\def\<{\langle}
\newcommand\hull{\mathrm{hull}}

\begin{document}
\mainmatter              % start of a contribution
%
\title{Non-additive measures for quantum probability?}
%
\titlerunning{Non-additive measures for quantum probability?}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Gabriele Carcassi \and Christine A. Aidala}
%
\authorrunning{Gabriele Carcassi and Christine A. Aidala} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
%\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
%Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\institute{University of Michigan, Ann Arbor, MI 48109, USA\\
\email{carcassi@umich.edu}
}

\maketitle              % typeset the title of the contribution

\begin{abstract}
	It is well-established that quantum probability does not follow classical Kolmogorov probability calculus. Various approaches have been developed to loosen the axioms, of which the use of signed measures is the most successful (e.g. the Wigner quasiprobability distribution). As part of our larger effort Assumptions of Physics, we have been considering the various roles of measures, which are used in physics not only for probability, but also to quantify the count of possible states and configurations. These measures play a crucial role in classical mechanics, as they effectively define its geometric structure. If one tries to construct a parallel in quantum mechanics, the measure to quantify the count of states turns out to be non-additive. The idea, then, is that the proper extension of probability calculus may require the use of non-additive measures, which is something that, to our knowledge, has not yet been explored.

	The purpose of this paper is to present the general idea and the open problems to an audience that is knowledgeable of the subject of non-additive set functions, though not necessarily in quantum physics, in the hope that it will spark helpful discussions. We will go through the motivation in simple terms, which stems from the link between the entropy $S$ of a uniform distribution $\rho_U$ and the logarithm of the measure $\mu$ associated to its support $U$ (i.e. $S(\rho_U) = \log (\mu(U))$). If one extends this notion to quantum mechanics, the associated measure $\mu$ is non-additive. We will explore some properties of this “quantum measure”, its reasonableness in terms of the physics, but its peculiarity on the math side. We will explore the need for a set of properties that can properly characterize the measure and a generalization of the Radon-Nikodym derivative to define a properly extended probability calculus that reduces to the standard additive one on sets of physically distinguishable cases (i.e. orthogonal measurement outcomes).

\keywords{non-additive set functions, quantum physics, statistical mechanics, information theory}
\end{abstract}
%
\section{Introduction}
%
It is well-established that quantum probability does not follow classical Kolmogorov probability calculus. Various approaches have been developed to loosen the axioms, of which the use of signed measures is the most successful (e.g. the Wigner quasiprobability distribution).\cite{gleason1957measures,groenewold1946principles,gudder2009quantum,hamhalter2003quantum,monchietti2023measure,moyal1949quantum,sorkin1994quantum,svozil2022extending} On the other hand, the use of non-additive measures (i.e. set functions) has received little attention, which is striking for two reasons. First, we have found that non-additive measures emerge quite naturally in quantum mechanics when looking for a quantum analogue of count of states\cite{aop-phys-QuantumRequiresNonAdditiveMeasures}, which plays a fundamental role in statistical mechanics. Second, there has been an increased use of quantum probability in non-classical decision making CITE, an area where non-additive measure theory is widely used.

The aim of this article is to introduce the problem for those that work on non-additive set functions but are not familiar with quantum mechanics. The overall goal is to find experts in the field that are interested in collaborating or helping. The general idea is that, in classical mechanics, one is provided with two additive measures, a Liouville measure to quantify states and a measure of probability, and the Radon-Nikodym derivative of the two returns the probability density. The goal is to find a quantum analogue for the two measures and the derivative, thus providing a full ``measure theoretic'' generalization.

\section{Measures in classical mechanics}

In classical mechanics, an ensemble is given by a probability measure $p$ over the Borel algebra of phase space $X$ (i.e. the symplectic manifold charted by the position $q^i$ and momentum $k_i$ of all particles of the system). The space of classical ensembles $\mathcal{E}_C$ (i.e. the space of probability measures over $X$) is a convex space as a convex combination $\lambda p_1 + (1 - \lambda) p_2$ with $\lambda \in [0,1]$ of two probability measures $p_1$ and $p_2$ is also a probability measure.

The space is also equipped with a measure $\mu(U) = \int_U \prod_i dq^i dk_i$ that quantifies states in a volume $U$. Given an ensemble $p$, the probability density $\rho = \frac{dp}{d\mu}$ is the Radon-Nikodym derivative and its entropy is $S(\rho) = - \int_X \rho \log \rho d\mu$.\footnote{The density and entropy must be computed with the Liouville measure $\mu$ or one does not recover the correct physics.} For a uniform distribution over a set $U$, $S(\rho_U) = \log \mu(U)$. Note that we will assume all logarithms to be in base 2, as is customary in information theory.

\section{Finding measures in quantum mechanics}

In quantum mechanics, the space of (pure) states $X$ is the projective space $P(\mathcal{H})$ of a Hilbert space $\mathcal{H}$. An ensemble (or mixed state) is given by a positive semi-definite Hermitian operator $\rho$ of trace one acting on the Hilbert space $\mathcal{H}$ of the system. The details of Hilbert spaces are not essential for our discussion; what is important is that the space of quantum ensembles $\mathcal{E}_Q$ is also a convex space as a convex combination $\lambda \rho_1 + (1 - \lambda) \rho_2$ with $\lambda \in [0,1]$ of two positive semi-definite Hermitian operators $\rho_1$ and $\rho_2$ of trace one is also a positive semi-definite Hermitian operator of trace one. A pure state is an extreme point of the convex set (i.e. an ensemble that cannot be further decomposed into a convex combination of other ensembles) and any ensemble $\rho$ can be expressed as a convex combination of countably many pure states. To each ensemble is associated an entropy $S(\rho) = - tr(\rho \log \rho)$.

The goal is to characterize a quantum ensemble $\rho$ as a set function $p$ over the Borel algebra of the pure states $X$ in the same way a classical ensemble is a probability measure over phase space. Additionally, we want a set function $\mu$ that quantifies quantum states, similarly to the classical Liouville measure, and to define the equivalent to the probability density $\frac{dp}{d\mu}$. Finally, we want to make sure that classical distributions (i.e. additive measures) are recovered for each quantum observable.\footnote{The idea is that the Borel algebra will contain the lattice of closed subspaces, and therefore we simply need to recover some key results on that sub-lattice.}

\subsection{First attempt for a quantum state-quantifying measure}

As a first attempt\cite{aop-phys-QuantumRequiresNonAdditiveMeasures}, we imposed the same relationship between measure and entropy one has in classical mechanics. That is, given a set $U \subseteq X$ of pure states, we consider the ensemble $\rho_U$ constructed with a uniform convex combination\footnote{The set of pure states $X$ is equipped with a measure invariant under unitary transformation, which can be used to define uniform convex combinations over a Borel set of $X$.}, and we define $\mu(U) = 2^{S(\rho_U)}$.

Since the entropy of a pure state is zero in quantum mechanics, we have $\mu(\{\psi\}) = 1$. For a set of two pure states $U = \{ \psi, \phi \}$, we have 
\begin{equation}
	S(\rho_U) = - \frac{1+\sqrt{p}}{2} \log \frac{1+\sqrt{p}}{2} 
	- \frac{1-\sqrt{p}}{2} \log \frac{1-\sqrt{p}}{2}.
\end{equation}
where $p=|\<\psi | \phi \>|^2$ is given by the square of the inner product, and can be physically understood as the probability of measuring one state given that the other was prepared. Since $0 \leq p \leq 1$, we have $1 \leq S(\rho_U) \leq 2$, which means that $\mu$ is not additive.

Physically, the measure is additive only when the two states are perfectly distinguishable through a measurement or, equivalently, when they can be prepared in the same physical conditions. We can therefore understand $\mu$ as counting states at-all-else-being-equal, and we have arguments as to why this is physically appropriate.

There is, however, an additional property of $\mu$ that make it less appealing: it is non-monotone. That is, the entropy of a uniform distribution over three states can have lower entropy than the uniform distribution of a subset of two. On physical grounds, saying that a bigger set of states has fewer states at-all-else-being-equal does not seem right. On mathematical grounds, as we learned from discussing the topic with experts in non-additive measure theory, generalizations such as the Choquet integral require monotone measures. This led us to revise the approach.

\subsection{Second attempt for quantum measures}

Our second approach, which is still being finalized\cite{aop-book}, defines $\mu(U) = \sup(2^{S(\hull(U))})$ as the supremum of the entropy reachable with a convex combination of the elements of the set.\footnote{In classical mechanics, this definition recovers $\mu(U) = 2^{S(\rho_U)}$ as the uniform distribution returns the maximum entropy.} If $U$ is a set of one or two pure states, the measure is the same as what we already computed. Therefore the main findings of the previous work carry over. For bigger sets, the measure is monotone because the supremum as well as the hull are monotone functions. Moreover, $\mu$ can be shown to be sub-additive.

Given a mixed state $\rho$, we can define $p_{\rho}(U) = \sup(\{ \lambda \in [0,1] \, | \, \exists \, \rho_1 \in \hull(U), \rho_2 \in \mathcal{E}_Q \; s.t. \;  \rho = \lambda \rho_1 + (1-\lambda) \rho_2 \})$ as the biggest fraction of $\rho$ that can be expressed as a convex combination of $U$. This set function can also be shown to be monotone and sub-additive.

When $\mu$ and $p_\rho$ are restricted to the lattice of closed subspaces, they both become additive on orthogonal subspaces. With this insight we should be able to recover classical probability for quantum measurements.

\section{Open questions}

While we have identified possible candidates for our measures, there are many questions that remain open. First of all, what is the proper notion of derivative, and therefore of integral, that we should be using? Ideally, for every quantum observable $O$ we would like a function $o : X \to \mathbb{R}$ of the pure states so that integrating that function recovers the expectation value for a given mixed state. That is, $tr(\rho O) = \int_X o \frac{dp_\rho}{d\mu} d\mu$. Similarly, we would like to express the entropy as an integral. It is not clear to us whether these features are possible or not.

Additionally, it would be interesting to understand whether these non-additive set functions have any relationship to the sorts of problems for which non-additive set functions are used in decision theory or other areas. We look forward to productive discussions in these areas.

\section{Acknowledgements}

This paper was made possible through the support of grant \#62847 from the John Templeton Foundation. It is part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.  We thank Vicenc Torra and Michel Grabisch for useful pointers and discussion.

%
% ---- Bibliography ----
%
\bibliographystyle{spmpsci}
\bibliography{bibliography}
\end{document}

### 2024-05-20 Proceedings/2024-IPMU/carcassi-quantummeasure.tex
IPMU2024: adding last citation

@@
 First, we have found that non-additive measures emerge quite naturally in quantum mechanics when looking for a quantum analogue of count of states\cite{aop-phys-QuantumRequiresNonAdditiveMeasures}, which plays a fundamental role in statistical mechanics.
-Second, there has been an increased use of quantum probability in non-classical decision making CITE, an area where non-additive measure theory is widely used.
+Second, there has been an increased use of quantum probability in non-classical decision making\cite{quantumrev2023}, an area where non-additive measure theory is widely used.
 

### 2024-08-27 Articles/2023-HilbertQuantum/Submissions/Quantum Studies/HilbertQuantum.tex
Quantum Studies submission; FoP cover letter
HilbertQuantum: change of language after editor comments

--- 2024-01-23 Articles/2023-HilbertQuantum/HilbertQuantum.tex
@@
 \begin{abstract}
-We show that Hilbert spaces should not be considered the ``correct'' spaces to represent quantum states mathematically.
+We show that Hilbert spaces are not suitable to represent quantum states mathematically, in the sense that they require properties that are untenable by physical entities.
 We first prove that the requirements posited by complex inner product spaces are physically justified.
 We then show that completeness in the infinite-dimensional case requires the inclusion of states with infinite expectations, coordinate transformations that take finite expectations to infinite ones and vice-versa, and time evolutions that transform finite expectations to infinite ones in finite time.
-This makes Hilbert spaces physically unsound as they model a potential infinity as an actual infinity.
-We suspect that at least some problems in quantum theory related to infinities may be ultimately caused by the wrong space being used.
-We strongly believe a better solution can be found, and we look at Schwartz spaces for inspiration, as, among other things, they guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are closed under Fourier transform.
+This means we should be wary of using Hilbert spaces to represent quantum states as they turn a potential infinity into an actual infinity.
+The main point of the paper, then, is to raise awareness that results that rely on the completeness of Hilbert spaces may not be physically significant.
+While we do not claim to know what a physically more appropriate closure should be, we note that Schwartz spaces, among other things, guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are closed under Fourier transform.
 \end{abstract}
@@
 
-In an effort to find a precise physical justification, we realized that no such justification can exist as infinite-dimensional Hilbert spaces hopelessly contain unphysical elements.
+In an effort to find a precise physical justification, we realized that no such justification can exist as infinite-dimensional Hilbert spaces always contain elements that cannot have physical counterparts.
 To put it simply, they must include states that have infinite expectation values for position, momentum, energy and so on.
@@
 Clearly, this does not make physical sense.
-When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that the observed value can be arbitrarily large, not that we can literally observe and infinite value.\footnote{This is the difference between potential infinity and actual infinity.}
+When we posit that a quantity, like position, has an infinite range of possible values, we simply mean that the observed value can be arbitrarily large, not that we can literally observe an infinite value.\footnote{This is the difference between potential infinity and actual infinity.}
 
@@
 For example, changes of basis and changes of variables are unitary transformations.
-The second reason is that time evolution for an isolated system is represented by a unitary transformation.
+This is already a problem: the space of wave-functions for $n$ particles is the Hilbert space $L^2(\mathbb{R}^{3n})$.
+For any value of $n$, these spaces have a countable basis (i.e.
+infinite dimensional and separable) and are therefore all mathematically equivalent.
+That is, the state space for one particle is the same as the state space for, say, a billion particles.
+We can therefore theoretically construct a quantum channel that encodes the information of a billion particles into a single one.
+Note that this is not possible in classical mechanics: the state space for $n$ classical particles is $\mathbb{R}^{6n}$.
+For different $n$, these spaces are not even topologically equivalent.
+
+The second reason for the importance of unitary transformation is that they represent possible time evolution for an isolated system.
 The issue is that unitary transformations in a Hilbert space can map states with finite expectation values to those with infinite ones, leading to obvious problems.
@@
 
-\section{On the physicality of Schwartz spaces}
+\section{On the possible physicality of Schwartz spaces}
 
@@
 A reasonable alternative would be to require the expectation of all polynomials of position and momentum to be finite.
-In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m - P^mX^n)$ to be finite.
+In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m + P^mX^n)$ to be finite.
 This restricts our space to those functions that are infinitely smooth (i.e.
@@
 \item it is complete with respect to the expectation values - a sequence of functions for which all expectation values of all polynomials of position and momentum converge will converge in the Schwartz space
+\item the Schwartz spaces over $\mathbb{R}^{n}$ are topologically different if $n$ is different
 \item its closure over all Cauchy sequences recovers the Hilbert space - the only thing missing are the Cauchy sequences for which the limit of at least one expectation is not well defined
@@
 \end{itemize}
-These properties make the Schwartz space a much more reasonable candidate to capture the physics.\cite{albert2022bosonic} The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} To again be clear, we are not proposing this as a solution.
-We are merely pointing out that alternatives do exist and should be fully explored.\footnote{Note that $C^*$-algebras model observables and therefore do not solve the problem at hand of modeling states.}
+These properties make the Schwartz space a much more reasonable candidate to capture the physics.\cite{albert2022bosonic} The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} However, it does not seem to us that this is a final and general answer.
+It is not clear to us how Schwartz spaces are generalized in the case where the topology of the underlying space is not $\mathbb{R}^n$.
+Moreover, something more general will be needed in the case of indistinguishable particles of variable number, which is handled by the Fock space in the standard treatment.
+However, we do think that the properties of Schwartz spaces may provide guidance for a possible comprehensive solution.\footnote{Note that $C^*$-algebras model observables and therefore do not solve the problem at hand of modeling states.}
 
@@
 
+\section*{Data Availability Statement}
+Not applicable.
+
+\section*{Conflict of interest}
+On behalf of all authors, the corresponding author states that there is no conflict of interest.
+
 \bibliography{bibliography}
@@
 This appendix presents the same arguments in the article's main body, but expanded with full mathematical detail.
+The figure and the derivations in the Appendix are courtesy of the Assumptions of Physics project.
+\cite{aop-book}
 
@@
 
+Also note that Schwartz spaces over $\mathbb{R}^n$ are topologically different if $n$ is different.
+This means that Schwartz spaces can distinguish the number of the degrees of freedom at the topological level.
+This is in line with what happens in classical mechanics, where the phase space $\mathbb{R}^{2n}$ charted by position and momentum is topologically different for different values of $n$.
+Additionally, the completion of those Schwartz spaces for any $n$ is the same Hilbert space.
+This means that the completion actually loses physical information about the space, which again shows that the physical content is actually in the Schwartz space.
+
 The Schwartz space also has another important property: it is closed under the Fourier transform.
@@
 Second, we would need to show the importance of expectation values of polynomials from experimental considerations.\footnote{The lack of a clear classical parallel for the moment problem restricted to pure states may make that argument difficult to construct.} It is not clear to us that that's the case, so it may be that the justification would have to proceed in a different, but ultimately equivalent, way.
+
+Additionally, Schwartz spaces would be physical only if we are modeling a fixed number of degrees of freedom over Euclidean space.
+We would still need a more general definition that would apply to all cases and would reduce to Schwartz spaces in the above case.
 

### 2024-08-27 Proceedings/2023-QIP/response1-reviewer2.txt
QIP: archiving reviewer responses

We thank the reviewer for the positive and interesting feedback. We really appreciate they feel we were able to present a highly technical and abstract subject in a clear and accessible way. Many comments go beyond the scope of the present paper and would certainly invite interesting discussion. If the reviewer is inclined, after the review process is completed, we would be delighted to continue the conversation.

– The authors present this work as part of a major research project
they call reverse physics, "which, starting from the laws, aims to
find a minimal set of physical assumptions needed to rederive them".
Trying to contextualize it, I immediately thought of axiomatization.
Indeed, axiomatizations aims at exactly that, reducing a body of
theorems to a minimal set that contains them within its logical
implications, without redundancy or contradiction. It has been
introduced for mathematical theories, but should quite as well apply
to sufficiently formalized physical theory. So, what's the
difference? It it lies in the working direction, i.e., here,
extracting axioms from a given set of laws (analogous to inverse
scattering), it would be interesting to whether there exist meta-
theorems in logics indicating the prospects of this process, for
example, by determining the number of required axioms as a kind of
dimension of the space of propositions spanned by the laws.

It would probably take a full paper to delve into these issues. First, it should be clear that mere axiomatization does not provide much insight into a physical theory. After all, the postulates of quantum mechanics can be considered axioms of the theory. In our mind, the issue is that the starting points, whatever we call them, should ideally be understandable in clear physics terms, which presents some additional problems.

The first issue is the transition between the informal physical definitions (which are needed as they connect to experiments) to the formal language of the theory. During discussions, at least two philosophers of science argued that this was one of the key new elements in Galilean thinking: he knowingly studied idealized pendulums or falling bodies neglecting friction. Therefore there are considerations to be had even before the formal system is set up. We suspect, in fact, that some sort of idealization/simplification is a prerequisite to any formal physical theory. A clear understanding of how the transition from the informal side to the formal side works is part of our goal.

The second issue is that the same assumption can be implemented formally in different ways. For example, the assumptions of determinism and reversibility can be implemented by requiring equal number of states to be mapped forward and backward in time; information on the initial state to correspond to equal information on the final state; the uncertainty for a peaked distribution to be conserved during evolution and so on. Therefore, you have one physical assumption that maps to multiple mathematical translations (i.e. multiple axioms) which can be proven to be equivalent. This is like the axiom of choice being equivalent to Zorn's lemma or the well-ordering theorem. Studying equivalences between these starting points is more important to us than any particular choice of starting point.

It was actually another philosopher that told us that our approach was similar to how reverse mathematics operates, and therefore we adopted the name. Some of these issues are mentioned in our paper on reverse physics. Some not, as we have been gradually becoming more aware of them. Ideally, we would work with someone from the philosophy side to give a more clear conceptual framework to the reverse physics approach.

– Concerning the specific case of quantum mechanics, I wwonder whether
the authors are aware of other current projects which perhaps are very
distinct as to their conceptual framework, but very close in spirit
to the present work. In particular, there exist research activities
aiming at an axiomatization of quantum mechanics that includes basic
postulates on the information content or capacity of quantum systems
(for a selection, not quite up to date, see my refs. [1-5] below).
Mentioning these efforts would be instructive for readers and relevant
regarding originality.

We were aware of Hardy and Chiribella et al., with the second considered closer as they are not constrained to quantum mechanics. It is probably best to refer to the reverse physics paper, as this was written to give a sense of the difference between our approach and the others. We added a footnote:

"To see how the reverse physics approach compares and contrasts with respect to other works that aim to find physically motivated starting points, for example in the context of quantum theory,[11-15] we refer to [10]."

– Including also mixed states appears consequent in view of the
objectives of the project. In quantum mechanics, however, mixed states
occur only in open systems, by masking external components of
the system which, taken into account, would render the total system
closed and thus in a pure state. They therefore implicitly involve
additional states, even a macroscopic environment if, e.g.,
measurements are involved. I am not sure whether an attempt to count
states makes sense if the very borders of the system are not well
defined.

The reviewer here is touching on another point that we are also working on: what are the assumptions/requirements needed to be able to define and characterize a physical system? Defining what happens at the boundary is indeeded crucial, and, in some sense, we believe that systems can only be defined if there are processes under which the system is, at least to some good approximation, isolated. These are details we will need to address in future works. We have added the following paragraph in the future work section:

"Stepping back from the mathematical problem, there are also more conceptual issues that need to be understood. For example, what are the operational requirements to be able to quantify the number of states? It may turn out that the ability to operationally define and manipulate a system already provides all that is needed to define how to quantify states."

That being said, we do believe that mixed states are a core concept. As for the objection raised, to our understanding, it is only improper mixtures (i.e. those describing a subsystem of an entangled system) that require another system. Proper mixtures (i.e. those that cannot be understood as coming from an entangled state) can arise in different ways, the most obvious being due to ignorance on the preparation. Moreover, if one can only perform experiments on a particular system, proper and improper mixtures are indistiguishable. It is only when we are able to define two independent systems that we can show that the mixture on one was improper. Therefore, when defining a single system, we can't technically make the distinction.

Additionally, all that we can prepare and measure in practice are mixtures. We can never perfectly prepare or measure pure states. Even if we prepare an electron with spin up, there will be an uncertainty in the direction. At any rate, these are issues that are important, but go outside the scope of this paper.

– Evidently, the non-additivity of quantum state measures is
intimately related to the possibility of finite overlaps of quantum
states, which in turn is a consequence of how quantum mechanics
defines distances between Hilbert-space vectors. This indicates a
relation between metric and measure on (hyper)spheres, which could add
an important argument to the paper.

The reviewer is again touching on another point we are already exploring. Both the spaces of classical and quantum mixtures are metric spaces under the Fisher-Rao metric. This can be understood as the infinitesimal version of the Jensen-Shannon divergence, which essentially tells us how the entropy increases under mixture. If two mixtures are non-overlapping, the JS divergence has maximum increase, both in classical and quantum mechanics. It should be possible to reframe this link in terms of state counting.

Given that it would bring a whole set of new concepts to the paper, which would take significant space and add complexity, we had decided not to include this insight. Though it is clearly a target for future work.

– The authors base part of their reasoning on a comparison of
classical with quantum measures in the particular case of spherical
state spaces. I doubt that this analogy is correct. In the classical
case, this space could represent, for instance, the orientation of an
angular momentum in three-dimensional space. In quantum mechanics, it
refers to the Hilbert space of a two-state system. However, these
spaces are not compatible. The abstract spherical Hilbert space of a
two-state system\ has nothing to do with angle coordinates three-
dimensional position space. Moreover, if mixed states come into play,
say represented by a Bloch sphere, they do not correspond to
probability densities on the surface, as the authors insinuate, but to
points inside the sphere, i.e., to Bloch vectors with a radius less
than one.

Note that we are looking only for a mathematical comparison of the measures. Clearly the state spaces are different and the spaces of mixtures are not comparable. However, note that the Bloch sphere is the state space for a spin one-half system, which is the orientation of an (intrinsic) angular momentum in three dimensional space. Therefore the comparison between the two is not that stretched. We added a paragraph to clarify:

"Note that the comparison is only at the mathematical level, so the only requirement is to find a mathematical space that can be used to represent states in the three cases. The sphere is used in classical mechanics to define the orientation of an angular momentum in three dimensions; the manifold is symplectic and allows both a Liouville measure and a counting measure. The (Bloch) sphere is used in quantum mechanics to represent the states of a two-state system, which can also represent the orientation of a spin one-half particle, and therefore it also allows a quantized measure."

– The argument, put forward in section 5, that a quantum state measure
attributing finite counts to finite continuous regions must be bounded
from below, is evident. An immediate consequence is the concept of
Planck cells, i.e., regions in phase space corresponding to that lower
bound. This direct relationship to one of the basic traits of quantum
mechanics might render this reasoning even more convincing. A further
consequence, distinguishing quantum mechanics from a direct rigid
discretization of space and time, notably to limit the information
content of these quantities, is that quantization fixes the size, but
in no way the shape of Planck cells.

The reviewer is again touching on one of our related research directions. In the reverse physics paper, we use the lower bound on entropy (which is the same as a lower bound on the state count) to motivate a classical analogue of the uncertainty principle. The issue of Planck cells, in fact, is that even knowing that there are cells already implies that one can define a boundary between them in a sharp way, but that definition would imply the ability to resolve the space between cells, which is a contradiction. We suspect that quantum mechanics is essentially the only way to implement a zero entropy lower bound in a consistent way. This would be equivalent to showing the uniqueness of the quantized measure. So, indeed, this is all related.

While, as the reviewer notes, there is a strong link between these ideas, we decided not to include them as we found that it may overwhelm some readers. One crazy idea at a time! :-)

– An interesting historical aspect is that quantization as a
consequence of a boundedness of entropy has been anticipated clearly
by Planck, in his derivation of a finite quantum of action, and even
earlier by Gibbs, proposing his paradox, and by Boltzmann himself,
referring to discrete "complexions" in his definition of entropy.

We were aware of some of Planck's work (a couple of philosophers made us aware of it) but we had no idea of similar efforts from Boltzmann. This is another topic of potential collaboration with someone more knowledgable with that literature: in a way, we are trying to understand how those arguments could be made to work to derive quantum mechanics.

– Sections 2 and 3 have the same title ("Measure theory and
quantifying states"). Certainly an unwanted mishap.

We are extremely grateful to the reviewer to have caught this! We modified the title of section 3 to "Entropy and state quantification".

### 2024-08-27 Proceedings/2024-IPMU/carcassi-quantummeasure.tex
IPMU: adding IPMU proceedings final version

--- 2024-05-18 Proceedings/2024-IPMU/carcassi-quantummeasure.tex
@@
 First, we have found that non-additive measures emerge quite naturally in quantum mechanics when looking for a quantum analogue of count of states\cite{aop-phys-QuantumRequiresNonAdditiveMeasures}, which plays a fundamental role in statistical mechanics.
-Second, there has been an increased use of quantum probability in non-classical decision making CITE, an area where non-additive measure theory is widely used.
+Second, there has been an increased use of quantum probability in non-classical decision making (see e.g.
+Ref.
+\cite{quantumrev2023}), an area where non-additive measure theory is widely used.
 
@@
 
-Our second approach, which is still being finalized\cite{aop-book}, defines $\mu(U) = \sup(2^{S(\hull(U))})$ as the supremum of the entropy reachable with a convex combination of the elements of the set.\footnote{In classical mechanics, this definition recovers $\mu(U) = 2^{S(\rho_U)}$ as the uniform distribution returns the maximum entropy.} If $U$ is a set of one or two pure states, the measure is the same as what we already computed.
+Our second approach, still being finalized\cite{aop-book}, defines $\mu(U) = \sup(2^{S(\hull(U))})$ as the supremum of the entropy reachable with a convex combination of the elements of the set.\footnote{In classical mechanics, this definition recovers $\mu(U) = 2^{S(\rho_U)}$ as the uniform distribution returns the maximum entropy.} If $U$ is a set of one or two pure states, the measure is the same as what we already computed.
 Therefore the main findings of the previous work carry over.

### 2024-08-27 Proceedings/2024-IPMU/editor/editor.tex
IPMU: adding IPMU proceedings final version

% This is EDITOR.TEX the demonstration file of
% the LaTeX macro package from Springer Nature
% for Proceedings volumes using the LaTeX document
% class SVProc - version as of 22. July 2016
%
\documentclass{SVProc}
%
\usepackage{makeidx}  % allows for indexgeneration
\makeindex
%
% to typeset URLs, URIs, and DOIs
\usepackage{url}
\def\UrlFont{\rmfamily}
%
\begin{document}
%
\frontmatter          % for the preliminaries
%
\pagestyle{headings}  % switches on printing of running heads
%%%\addtocmark{Hamiltonian Mechanics} % additional mark in the TOC
%
\chapter*{Preface}
%
This textbook is intended for use by students of physics, physical
chemistry, and theoretical chemistry. The reader is presumed to have a
basic knowledge of atomic and quantum physics at the level provided, for
example, by the first few chapters in our book {\it The Physics of Atoms
and Quanta}. The student of physics will find here material which should
be included in the basic education of every physicist. This book should
furthermore allow students to acquire an appreciation of the breadth and
variety within the field of molecular physics and its future as a
fascinating area of research.

\vspace{1cm}
\begin{flushright}\noindent
June 2016\hfill Walter Olthoff\\
Program Chair\\
ECOOP 2016
\end{flushright}
%
\chapter*{Organization}
ECOOP 2016 is organized by the department of Computer Science, Univeristy
of \AA rhus and AITO (association Internationa pour les Technologie
Object) in cooperation with ACM/SIGPLAN.
%
\section*{Executive Commitee}
\begin{tabular}{@{}p{5cm}@{}p{7.2cm}@{}}
Conference Chair:&Ole Lehrmann Madsen (\AA rhus University, DK)\\
Program Chair:   &Walter Olthoff (DFKI GmbH, Germany)\\
Organizing Chair:&J\o rgen Lindskov Knudsen (\AA rhus University, DK)\\
Tutorials:&Birger M\o ller-Pedersen\hfil\break
(Norwegian Computing Center, Norway)\\
Workshops:&Eric Jul (University of Kopenhagen, Denmark)\\
Panels:&Boris Magnusson (Lund University, Sweden)\\
Exhibition:&Elmer Sandvad (\AA rhus University, DK)\\
Demonstrations:&Kurt N\o rdmark (\AA rhus University, DK)
\end{tabular}
%
\section*{Program Commitee}
\begin{tabular}{@{}p{5cm}@{}p{7.2cm}@{}}
Conference Chair:&Ole Lehrmann Madsen (\AA rhus University, DK)\\
Program Chair:   &Walter Olthoff (DFKI GmbH, Germany)\\
Organizing Chair:&J\o rgen Lindskov Knudsen (\AA rhus University, DK)\\
Tutorials:&Birger M\o ller-Pedersen\hfil\break
(Norwegian Computing Center, Norway)\\
Workshops:&Eric Jul (University of Kopenhagen, Denmark)\\
Panels:&Boris Magnusson (Lund University, Sweden)\\
Exhibition:&Elmer Sandvad (\AA rhus University, DK)\\
Demonstrations:&Kurt N\o rdmark (\AA rhus University, DK)
\end{tabular}
%
\begin{multicols}{3}[\section*{Referees}]
V.~Andreev\\
B\"arwolff\\
E.~Barrelet\\
H.P.~Beck\\
G.~Bernardi\\
E.~Binder\\
P.C.~Bosetti\\
Braunschweig\\
F.W.~B\"usser\\
T.~Carli\\
A.B.~Clegg\\
G.~Cozzika\\
S.~Dagoret\\
Del~Buono\\
P.~Dingus\\
H.~Duhm\\
J.~Ebert\\
S.~Eichenberger\\
R.J.~Ellison\\
Feltesse\\
W.~Flauger\\
A.~Fomenko\\
G.~Franke\\
J.~Garvey\\
M.~Gennis\\
L.~Goerlich\\
P.~Goritchev\\
H.~Greif\\
E.M.~Hanlon\\
R.~Haydar\\
R.C.W.~Henderso\\
P.~Hill\\
H.~Hufnagel\\
A.~Jacholkowska\\
Johannsen\\
S.~Kasarian\\
I.R.~Kenyon\\
C.~Kleinwort\\
T.~K\"ohler\\
S.D.~Kolya\\
P.~Kostka\\
U.~Kr\"uger\\
J.~Kurzh\"ofer\\
M.P.J.~Landon\\
A.~Lebedev\\
Ch.~Ley\\
F.~Linsel\\
H.~Lohmand\\
Martin\\
S.~Masson\\
K.~Meier\\
C.A.~Meyer\\
S.~Mikocki\\
J.V.~Morris\\
B.~Naroska\\
Nguyen\\
U.~Obrock\\
G.D.~Patel\\
Ch.~Pichler\\
S.~Prell\\
F.~Raupach\\
V.~Riech\\
P.~Robmann\\
N.~Sahlmann\\
P.~Schleper\\
Sch\"oning\\
B.~Schwab\\
A.~Semenov\\
G.~Siegmon\\
J.R.~Smith\\
M.~Steenbock\\
U.~Straumann\\
C.~Thiebaux\\
P.~Van~Esch\\
from Yerevan Ph\\
L.R.~West\\
G.-G.~Winter\\
T.P.~Yiou\\
M.~Zimmer\end{multicols}
%
\section*{Sponsoring Institutions}
%
V. Meyer Inc., Reading, MA, USA\\
The Hofmann-International Company, San Louis Obispo, CA, USA\\
Kramer Industries, Heidelberg, Germany
%
\tableofcontents
%
\mainmatter              % start of the contributions
%
\part{Hamiltonian Mechanics}
%
\title{Hamiltonian Mechanics unter besonderer Ber\"ucksichtigung der
h\"ohereren Lehranstalten}
%
\titlerunning{Hamiltonian Mechanics}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{Ivar Ekeland\inst{1} \and Roger Temam\inst{2}
Jeffrey Dean \and David Grove \and Craig Chambers \and Kim~B.~Bruce \and
Elsa Bertino}
%
\authorrunning{Ivar Ekeland et al.} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{Ivar Ekeland, Roger Temam, Jeffrey Dean, David Grove,
Craig Chambers, Kim B. Bruce, and Elisa Bertino}
%
\index{Ekeland, I.}
\index{Temam, R.}
\index{Dean, J.}
\index{Grove, D.}
\index{Chambers, C.}
\index{Kim, B.}
\index{Bertino, E.}
%
\institute{Princeton University, Princeton NJ 08544, USA,\\
\email{I.Ekeland@princeton.edu},\\ WWW home page:
\texttt{http://users/\homedir iekeland/web/welcome.html}
\and
Universit\'{e} de Paris-Sud,
Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
F-91405 Orsay Cedex, France}

\maketitle              % typeset the title of the contribution

\begin{abstract}
The abstract should summarize the contents of the paper
using at least 70 and at most 150 words. It will be set in 9-point
font size and be inset 1.0 cm from the right and left margins.
There will be two blank lines before and after the Abstract. \dots
\keywords{computational geometry, graph theory, Hamilton cycles}
\end{abstract}
%
\section{Fixed-Period Problems: The Sublinear Case}
%
With this chapter, the preliminaries are over, and we begin the search
for periodic solutions to Hamiltonian systems. All this will be done in
the convex case; that is, we shall study the boundary-value problem
\begin{eqnarray*}
  \dot{x}&=&JH' (t,x)\\
  x(0) &=& x(T)
\end{eqnarray*}
with $H(t,\cdot)$ a convex function of $x$, going to $+\infty$ when
$\left\|x\right\| \to \infty$.

%
\subsection{Autonomous Systems}
%
In this section, we will consider the case when the Hamiltonian $H(x)$
is autonomous. For the sake of simplicity, we shall also assume that it
is $C^{1}$.

We shall first consider the question of nontriviality, within the
general framework of
$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
the second subsection, we shall look into the special case when $H$ is
$\left(0,b_{\infty}\right)$-subquadratic,
and we shall try to derive additional information.
%
\subsubsection{The General Case: Nontriviality.}
%
We assume that $H$ is
$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
with $B_{\infty}-A_{\infty}$ positive definite. Set:
\begin{eqnarray}
\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
  J \frac{d}{dt} +A_{\infty}\ .
\end{eqnarray}

Theorem~\ref{ghou:pre} tells us that if $\lambda +\gamma < 0$, the
boundary-value problem:
\begin{equation}
\begin{array}{rcl}
  \dot{x}&=&JH' (x)\\
  x(0)&=&x (T)
\end{array}
\end{equation}
has at least one solution
$\overline{x}$, which is found by minimizing the dual
action functional:
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
\end{equation}
on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
with finite codimension. Here
\begin{equation}
  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
\end{equation}
is a convex function, and
\begin{equation}
  N(x) \le \frac{1}{2}
  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
  + c\ \ \ \forall x\ .
\end{equation}

%
\begin{proposition}
Assume $H'(0)=0$ and $ H(0)=0$. Set:
\begin{equation}
  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
  \label{eq:one}
\end{equation}

If $\gamma < - \lambda < \delta$,
the solution $\overline{u}$ is non-zero:
\begin{equation}
  \overline{x} (t) \ne 0\ \ \ \forall t\ .
\end{equation}
\end{proposition}
%
\begin{proof}
Condition (\ref{eq:one}) means that, for every
$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
\begin{equation}
  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
\end{equation}

It is an exercise in convex analysis, into which we shall not go, to
show that this implies that there is an $\eta > 0$ such that
\begin{equation}
  f\left\|x\right\| \le \eta
  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
  \left\|y\right\|^{2}\ .
  \label{eq:two}
\end{equation}

\begin{figure}
\vspace{2.5cm}
\caption{This is the caption of the figure displaying a white eagle and
a white horse on a snow field}
\end{figure}

Since $u_{1}$ is a smooth function, we will have
$\left\|hu_{1}\right\|_\infty \le \eta$
for $h$ small enough, and inequality (\ref{eq:two}) will hold,
yielding thereby:
\begin{equation}
  \psi (hu_{1}) \le \frac{h^{2}}{2}
  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
\end{equation}

If we choose $\delta '$ close enough to $\delta$, the quantity
$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
will be negative, and we end up with
\begin{equation}
  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
\end{equation}

On the other hand, we check directly that $\psi (0) = 0$. This shows
that 0 cannot be a minimizer of $\psi$, not even a local one.
So $\overline{u} \ne 0$ and
$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
\end{proof}
%
\begin{corollary}
Assume $H$ is $C^{2}$ and
$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
equilibria, that is, the solutions of $H' (\xi ) = 0$.
Denote by $\omega_{k}$
the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
\begin{equation}
  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
\end{equation}
If:
\begin{equation}
  \frac{T}{2\pi} b_{\infty} <
  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
  \frac{T}{2\pi}\omega
  \label{eq:three}
\end{equation}
then minimization of $\psi$ yields a non-constant $T$-periodic solution
$\overline{x}$.
\end{corollary}
%

We recall once more that by the integer part $E [\alpha ]$ of
$\alpha \in \bbbr$, we mean the $a\in \bbbz$
such that $a< \alpha \le a+1$. For instance,
if we take $a_{\infty} = 0$, Corollary 2 tells
us that $\overline{x}$ exists and is
non-constant provided that:

\begin{equation}
  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
\end{equation}
or
\begin{equation}
  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
  \label{eq:four}
\end{equation}

%
\begin{proof}
The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
largest negative eigenvalue $\lambda$ is given by
$\frac{2\pi}{T}k_{o} +a_{\infty}$,
where
\begin{equation}
  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
\end{equation}
Hence:
\begin{equation}
  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
\end{equation}

The condition $\gamma < -\lambda < \delta$ now becomes:
\begin{equation}
  b_{\infty} - a_{\infty} <
  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
\end{equation}
which is precisely condition (\ref{eq:three}).\qed
\end{proof}
%

\begin{lemma}
Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
\end{lemma}
%
\begin{proof}
We know that $\widetilde{x}$, or
$\widetilde{x} + \xi$ for some constant $\xi
\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
\begin{equation}
  \dot{x} = JH' (x)\ .
\end{equation}

There is no loss of generality in taking $\xi = 0$. So
$\psi (x) \ge \psi (\widetilde{x} )$
for all $\widetilde{x}$ in some neighbourhood of $x$ in
$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.

But this index is precisely the index
$i_{T} (\widetilde{x} )$ of the $T$-periodic
solution $\widetilde{x}$ over the interval
$(0,T)$, as defined in Sect.~2.6. So
\begin{equation}
  i_{T} (\widetilde{x} ) = 0\ .
  \label{eq:five}
\end{equation}

Now if $\widetilde{x}$ has a lower period, $T/k$ say,
we would have, by Corollary 31:
\begin{equation}
  i_{T} (\widetilde{x} ) =
  i_{kT/k}(\widetilde{x} ) \ge
  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
\end{equation}

This would contradict (\ref{eq:five}), and thus cannot happen.\qed
\end{proof}
%
\paragraph{Notes and Comments.}
The results in this section are a
refined version of \cite{smit:wat};
the minimality result of Proposition
14 was the first of its kind.

To understand the nontriviality conditions, such as the one in formula
(\ref{eq:four}), one may think of a one-parameter family
$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
of periodic solutions, $x_{T} (0) = x_{T} (T)$,
with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
which is the period of the linearized system at 0.

\begin{table}
\caption{This is the example table taken out of {\it The
\TeX{}book,} p.\,246}
\begin{center}
\begin{tabular}{r@{\quad}rl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt}
                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
\hline\rule{0pt}{12pt}
8000 B.C.  &     5,000,000& \\
  50 A.D.  &   200,000,000& \\
1650 A.D.  &   500,000,000& \\
1945 A.D.  & 2,300,000,000& \\
1980 A.D.  & 4,400,000,000& \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}
%
\begin{theorem} [Ghoussoub-Preiss]\label{ghou:pre}
Assume $H(t,x)$ is
$(0,\varepsilon )$-subquadratic at
infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
\begin{equation}
  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
\end{equation}
\begin{equation}
  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
\end{equation}
\begin{equation}
  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
\end{equation}
\begin{equation}
  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
\end{equation}

Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
$kT$-periodic solutions of the system
\begin{equation}
  \dot{x} = JH' (t,x)
\end{equation}
such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
\begin{equation}
  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
\end{equation}
\qed
\end{theorem}
%
\begin{example} [{{\rm External forcing}}]
Consider the system:
\begin{equation}
  \dot{x} = JH' (x) + f(t)
\end{equation}
where the Hamiltonian $H$ is
$\left(0,b_{\infty}\right)$-subquadratic, and the
forcing term is a distribution on the circle:
\begin{equation}
  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
\end{equation}
where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
\begin{equation}
  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
\end{equation}
where $\delta_{k}$ is the Dirac mass at $t= k$ and
$\xi \in \bbbr^{2n}$ is a
constant, fits the prescription. This means that the system
$\dot{x} = JH' (x)$ is being excited by a
series of identical shocks at interval $T$.
\end{example}
%
\begin{definition}
Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
operators in $\bbbr^{2n}$, depending continuously on
$t\in [0,T]$, such that
$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.

A Borelian function
$H: [0,T]\times \bbbr^{2n} \to \bbbr$
is called
$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
if there exists a function $N(t,x)$ such that:
\begin{equation}
  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
\end{equation}
\begin{equation}
  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
  {\rm is\ convex\ with\  respect\  to}\ \ x
\end{equation}
\begin{equation}
  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
\end{equation}
\begin{equation}
  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
\end{equation}

If $A_{\infty} (t) = a_{\infty} I$ and
$B_{\infty} (t) = b_{\infty} I$, with
$a_{\infty} \le b_{\infty} \in \bbbr$,
we shall say that $H$ is
$\left(a_{\infty},b_{\infty}\right)$-subquadratic
at infinity. As an example, the function
$\left\|x\right\|^{\alpha}$, with
$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
for every $\varepsilon > 0$. Similarly, the Hamiltonian
\begin{equation}
H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
\end{equation}
is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
Note that, if $k<0$, it is not convex.
\end{definition}
%

\paragraph{Notes and Comments.}
The first results on subharmonics were
obtained by Rabinowitz in \cite{fo:kes:nic:tue}, who showed the existence of
infinitely many subharmonics both in the subquadratic and superquadratic
case, with suitable growth conditions on $H'$. Again the duality
approach enabled Clarke and Ekeland in \cite{may:ehr:stein} to treat the
same problem in the convex-subquadratic case, with growth conditions on
$H$ only.

Recently, Michalek and Tarantello (see \cite{fost:kes} and \cite{czaj:fitz})
have obtained lower bound on the number of subharmonics of period $kT$,
based on symmetry considerations and on pinching estimates, as in
Sect.~5.2 of this article.

%
% ---- Bibliography ----
%
\begin{thebibliography}{6}
	%
	%
	\bibitem {smit:wat}
	Smith, T.F., Waterman, M.S.: Identification of common molecular subsequences.
	J. Mol. Biol. 147, 195?197 (1981). \url{doi:10.1016/0022-2836(81)90087-5}

	\bibitem {may:ehr:stein}
	May, P., Ehrlich, H.-C., Steinke, T.: ZIB structure prediction pipeline:
	composing a complex biological workflow through web services.
	In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006.
	LNCS, vol. 4128, pp. 1148?1158. Springer, Heidelberg (2006).
	\url{doi:10.1007/11823285_121}

	\bibitem {fost:kes}
	Foster, I., Kesselman, C.: The Grid: Blueprint for a New Computing Infrastructure.
	Morgan Kaufmann, San Francisco (1999)

	\bibitem {czaj:fitz}
	Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid information services
	for distributed resource sharing. In: 10th IEEE International Symposium
	on High Performance Distributed Computing, pp. 181?184. IEEE Press, New York (2001).
	\url{doi: 10.1109/HPDC.2001.945188}

	\bibitem {fo:kes:nic:tue}
	Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The physiology of the grid: an open grid services architecture for distributed systems integration. Technical report, Global Grid
	Forum (2002)

	\bibitem {onlyurl}
	National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}

\part{Theoretical Contributions} % yields also additional mark in the TOC

%
% second contribution with nearly identical text,
% slightly changed contribution head (all entries
% appear as defaults), and modified bibliography
%
\title{Hamiltonian Mechanics2}

\author{Ivar Ekeland\inst{1} \and Roger Temam\inst{2}}

\institute{Princeton University, Princeton NJ 08544, USA
\and
Universit\'{e} de Paris-Sud,
Laboratoire d'Analyse Num\'{e}rique, B\^{a}timent 425,\\
F-91405 Orsay Cedex, France}

%
\index{Ekeland, I.}
\index{Temam, R.}
\maketitle
%
% Modify the bibliography environment to call for the author-year
% system. This is done normally with the citeauthoryear option
% for a particular contribution.
\makeatletter
\renewenvironment{thebibliography}[1]
     {\section*{\refname}
      \small
      \list{}%
           {\settowidth\labelwidth{}%
            \leftmargin\parindent
            \itemindent=-\parindent
            \labelsep=\z@
            \if@openbib
              \advance\leftmargin\bibindent
              \itemindent -\bibindent
              \listparindent \itemindent
              \parsep \z@
            \fi
            \usecounter{enumiv}%
            \let\p@enumiv\@empty
            \renewcommand\theenumiv{}}%
      \if@openbib
        \renewcommand\newblock{\par}%
      \else
        \renewcommand\newblock{\hskip .11em \@plus.33em \@minus.07em}%
      \fi
      \sloppy\clubpenalty4000\widowpenalty4000%
      \sfcode`\.=\@m}
     {\def\@noitemerr
       {\@latex@warning{Empty `thebibliography' environment}}%
      \endlist}
      \def\@cite#1{#1}%
      \def\@lbibitem[#1]#2{\item[]\if@filesw
        {\def\protect##1{\string ##1\space}\immediate
      \write\@auxout{\string\bibcite{#2}{#1}}}\fi\ignorespaces}
\makeatother
%
\begin{abstract}
The abstract should summarize the contents of the paper
using at least 70 and at most 150 words. It will be set in 9-point
font size and be inset 1.0 cm from the right and left margins.
There will be two blank lines before and after the Abstract. \dots
\keywords{graph transformations, convex geometry, lattice computations,
convex polygons, triangulations, discrete geometry}
\end{abstract}
%
\section{Fixed-Period Problems: The Sublinear Case}
%
With this chapter, the preliminaries are over, and we begin the search
for periodic solutions to Hamiltonian systems. All this will be done in
the convex case; that is, we shall study the boundary-value problem
\begin{eqnarray*}
  \dot{x}&=&JH' (t,x)\\
  x(0) &=& x(T)
\end{eqnarray*}
with $H(t,\cdot)$ a convex function of $x$, going to $+\infty$ when
$\left\|x\right\| \to \infty$.

%
\subsection{Autonomous Systems}
%
In this section, we will consider the case when the Hamiltonian $H(x)$
is autonomous. For the sake of simplicity, we shall also assume that it
is $C^{1}$.

We shall first consider the question of nontriviality, within the
general framework of
$\left(A_{\infty},B_{\infty}\right)$-subquadratic Hamiltonians. In
the second subsection, we shall look into the special case when $H$ is
$\left(0,b_{\infty}\right)$-subquadratic,
and we shall try to derive additional information.
%
\subsubsection{The General Case: Nontriviality.}
%
We assume that $H$ is
$\left(A_{\infty},B_{\infty}\right)$-sub\-qua\-dra\-tic at infinity,
for some constant symmetric matrices $A_{\infty}$ and $B_{\infty}$,
with $B_{\infty}-A_{\infty}$ positive definite. Set:
\begin{eqnarray}
\gamma :&=&{\rm smallest\ eigenvalue\ of}\ \ B_{\infty} - A_{\infty} \\
  \lambda : &=& {\rm largest\ negative\ eigenvalue\ of}\ \
  J \frac{d}{dt} +A_{\infty}\ .
\end{eqnarray}

Theorem 21 tells us that if $\lambda +\gamma < 0$, the boundary-value
problem:
\begin{equation}
\begin{array}{rcl}
  \dot{x}&=&JH' (x)\\
  x(0)&=&x (T)
\end{array}
\end{equation}
has at least one solution
$\overline{x}$, which is found by minimizing the dual
action functional:
\begin{equation}
  \psi (u) = \int_{o}^{T} \left[\frac{1}{2}
  \left(\Lambda_{o}^{-1} u,u\right) + N^{\ast} (-u)\right] dt
\end{equation}
on the range of $\Lambda$, which is a subspace $R (\Lambda)_{L}^{2}$
with finite codimension. Here
\begin{equation}
  N(x) := H(x) - \frac{1}{2} \left(A_{\infty} x,x\right)
\end{equation}
is a convex function, and
\begin{equation}
  N(x) \le \frac{1}{2}
  \left(\left(B_{\infty} - A_{\infty}\right) x,x\right)
  + c\ \ \ \forall x\ .
\end{equation}

%
\begin{proposition}
Assume $H'(0)=0$ and $ H(0)=0$. Set:
\begin{equation}
  \delta := \liminf_{x\to 0} 2 N (x) \left\|x\right\|^{-2}\ .
  \label{2eq:one}
\end{equation}

If $\gamma < - \lambda < \delta$,
the solution $\overline{u}$ is non-zero:
\begin{equation}
  \overline{x} (t) \ne 0\ \ \ \forall t\ .
\end{equation}
\end{proposition}
%
\begin{proof}
Condition (\ref{2eq:one}) means that, for every
$\delta ' > \delta$, there is some $\varepsilon > 0$ such that
\begin{equation}
  \left\|x\right\| \le \varepsilon \Rightarrow N (x) \le
  \frac{\delta '}{2} \left\|x\right\|^{2}\ .
\end{equation}

It is an exercise in convex analysis, into which we shall not go, to
show that this implies that there is an $\eta > 0$ such that
\begin{equation}
  f\left\|x\right\| \le \eta
  \Rightarrow N^{\ast} (y) \le \frac{1}{2\delta '}
  \left\|y\right\|^{2}\ .
  \label{2eq:two}
\end{equation}

\begin{figure}
\vspace{2.5cm}
\caption{This is the caption of the figure displaying a white eagle and
a white horse on a snow field}
\end{figure}

Since $u_{1}$ is a smooth function, we will have
$\left\|hu_{1}\right\|_\infty \le \eta$
for $h$ small enough, and inequality (\ref{2eq:two}) will hold,
yielding thereby:
\begin{equation}
  \psi (hu_{1}) \le \frac{h^{2}}{2}
  \frac{1}{\lambda} \left\|u_{1} \right\|_{2}^{2} + \frac{h^{2}}{2}
  \frac{1}{\delta '} \left\|u_{1}\right\|^{2}\ .
\end{equation}

If we choose $\delta '$ close enough to $\delta$, the quantity
$\left(\frac{1}{\lambda} + \frac{1}{\delta '}\right)$
will be negative, and we end up with
\begin{equation}
  \psi (hu_{1}) < 0\ \ \ \ \ {\rm for}\ \ h\ne 0\ \ {\rm small}\ .
\end{equation}

On the other hand, we check directly that $\psi (0) = 0$. This shows
that 0 cannot be a minimizer of $\psi$, not even a local one.
So $\overline{u} \ne 0$ and
$\overline{u} \ne \Lambda_{o}^{-1} (0) = 0$. \qed
\end{proof}
%
\begin{corollary}
Assume $H$ is $C^{2}$ and
$\left(a_{\infty},b_{\infty}\right)$-subquadratic at infinity. Let
$\xi_{1},\allowbreak\dots,\allowbreak\xi_{N}$  be the
equilibria, that is, the solutions of $H' (\xi ) = 0$.
Denote by $\omega_{k}$
the smallest eigenvalue of $H'' \left(\xi_{k}\right)$, and set:
\begin{equation}
  \omega : = {\rm Min\,} \left\{\omega_{1},\dots,\omega_{k}\right\}\ .
\end{equation}
If:
\begin{equation}
  \frac{T}{2\pi} b_{\infty} <
  - E \left[- \frac{T}{2\pi}a_{\infty}\right] <
  \frac{T}{2\pi}\omega
  \label{2eq:three}
\end{equation}
then minimization of $\psi$ yields a non-constant $T$-periodic solution
$\overline{x}$.
\end{corollary}
%

We recall once more that by the integer part $E [\alpha ]$ of
$\alpha \in \bbbr$, we mean the $a\in \bbbz$
such that $a< \alpha \le a+1$. For instance,
if we take $a_{\infty} = 0$, Corollary 2 tells
us that $\overline{x}$ exists and is
non-constant provided that:

\begin{equation}
  \frac{T}{2\pi} b_{\infty} < 1 < \frac{T}{2\pi}
\end{equation}
or
\begin{equation}
  T\in \left(\frac{2\pi}{\omega},\frac{2\pi}{b_{\infty}}\right)\ .
  \label{2eq:four}
\end{equation}

%
\begin{proof}
The spectrum of $\Lambda$ is $\frac{2\pi}{T} \bbbz +a_{\infty}$. The
largest negative eigenvalue $\lambda$ is given by
$\frac{2\pi}{T}k_{o} +a_{\infty}$,
where
\begin{equation}
  \frac{2\pi}{T}k_{o} + a_{\infty} < 0
  \le \frac{2\pi}{T} (k_{o} +1) + a_{\infty}\ .
\end{equation}
Hence:
\begin{equation}
  k_{o} = E \left[- \frac{T}{2\pi} a_{\infty}\right] \ .
\end{equation}

The condition $\gamma < -\lambda < \delta$ now becomes:
\begin{equation}
  b_{\infty} - a_{\infty} <
  - \frac{2\pi}{T} k_{o} -a_{\infty} < \omega -a_{\infty}
\end{equation}
which is precisely condition (\ref{2eq:three}).\qed
\end{proof}
%

\begin{lemma}
Assume that $H$ is $C^{2}$ on $\bbbr^{2n} \setminus \{ 0\}$ and
that $H'' (x)$ is non-de\-gen\-er\-ate for any $x\ne 0$. Then any local
minimizer $\widetilde{x}$ of $\psi$ has minimal period $T$.
\end{lemma}
%
\begin{proof}
We know that $\widetilde{x}$, or
$\widetilde{x} + \xi$ for some constant $\xi
\in \bbbr^{2n}$, is a $T$-periodic solution of the Hamiltonian system:
\begin{equation}
  \dot{x} = JH' (x)\ .
\end{equation}

There is no loss of generality in taking $\xi = 0$. So
$\psi (x) \ge \psi (\widetilde{x} )$
for all $\widetilde{x}$ in some neighbourhood of $x$ in
$W^{1,2} \left(\bbbr / T\bbbz ; \bbbr^{2n}\right)$.

But this index is precisely the index
$i_{T} (\widetilde{x} )$ of the $T$-periodic
solution $\widetilde{x}$ over the interval
$(0,T)$, as defined in Sect.~2.6. So
\begin{equation}
  i_{T} (\widetilde{x} ) = 0\ .
  \label{2eq:five}
\end{equation}

Now if $\widetilde{x}$ has a lower period, $T/k$ say,
we would have, by Corollary 31:
\begin{equation}
  i_{T} (\widetilde{x} ) =
  i_{kT/k}(\widetilde{x} ) \ge
  ki_{T/k} (\widetilde{x} ) + k-1 \ge k-1 \ge 1\ .
\end{equation}

This would contradict (\ref{2eq:five}), and thus cannot happen.\qed
\end{proof}
%
\paragraph{Notes and Comments.}
The results in this section are a
refined version of \cite{2smit:wat};
the minimality result of Proposition
14 was the first of its kind.

To understand the nontriviality conditions, such as the one in formula
(\ref{2eq:four}), one may think of a one-parameter family
$x_{T}$, $T\in \left(2\pi\omega^{-1}, 2\pi b_{\infty}^{-1}\right)$
of periodic solutions, $x_{T} (0) = x_{T} (T)$,
with $x_{T}$ going away to infinity when $T\to 2\pi \omega^{-1}$,
which is the period of the linearized system at 0.

\begin{table}
\caption{This is the example table taken out of {\it The
\TeX{}book,} p.\,246}
\begin{center}
\begin{tabular}{r@{\quad}rl}
\hline
\multicolumn{1}{l}{\rule{0pt}{12pt}
                   Year}&\multicolumn{2}{l}{World population}\\[2pt]
\hline\rule{0pt}{12pt}
8000 B.C.  &     5,000,000& \\
  50 A.D.  &   200,000,000& \\
1650 A.D.  &   500,000,000& \\
1945 A.D.  & 2,300,000,000& \\
1980 A.D.  & 4,400,000,000& \\[2pt]
\hline
\end{tabular}
\end{center}
\end{table}
%
\begin{theorem} [Ghoussoub-Preiss]
Assume $H(t,x)$ is
$(0,\varepsilon )$-subquadratic at
infinity for all $\varepsilon > 0$, and $T$-periodic in $t$
\begin{equation}
  H (t,\cdot )\ \ \ \ \ {\rm is\ convex}\ \ \forall t
\end{equation}
\begin{equation}
  H (\cdot ,x)\ \ \ \ \ {\rm is}\ \ T{\rm -periodic}\ \ \forall x
\end{equation}
\begin{equation}
  H (t,x)\ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n (s)s^{-1}\to \infty\ \ {\rm as}\ \ s\to \infty
\end{equation}
\begin{equation}
  \forall \varepsilon > 0\ ,\ \ \ \exists c\ :\
  H(t,x) \le \frac{\varepsilon}{2}\left\|x\right\|^{2} + c\ .
\end{equation}

Assume also that $H$ is $C^{2}$, and $H'' (t,x)$ is positive definite
everywhere. Then there is a sequence $x_{k}$, $k\in \bbbn$, of
$kT$-periodic solutions of the system
\begin{equation}
  \dot{x} = JH' (t,x)
\end{equation}
such that, for every $k\in \bbbn$, there is some $p_{o}\in\bbbn$ with:
\begin{equation}
  p\ge p_{o}\Rightarrow x_{pk} \ne x_{k}\ .
\end{equation}
\qed
\end{theorem}
%
\begin{example} [{{\rm External forcing}}]
Consider the system:
\begin{equation}
  \dot{x} = JH' (x) + f(t)
\end{equation}
where the Hamiltonian $H$ is
$\left(0,b_{\infty}\right)$-subquadratic, and the
forcing term is a distribution on the circle:
\begin{equation}
  f = \frac{d}{dt} F + f_{o}\ \ \ \ \
  {\rm with}\ \ F\in L^{2} \left(\bbbr / T\bbbz; \bbbr^{2n}\right)\ ,
\end{equation}
where $f_{o} : = T^{-1}\int_{o}^{T} f (t) dt$. For instance,
\begin{equation}
  f (t) = \sum_{k\in \bbbn} \delta_{k} \xi\ ,
\end{equation}
where $\delta_{k}$ is the Dirac mass at $t= k$ and
$\xi \in \bbbr^{2n}$ is a
constant, fits the prescription. This means that the system
$\dot{x} = JH' (x)$ is being excited by a
series of identical shocks at interval $T$.
\end{example}
%
\begin{definition}
Let $A_{\infty} (t)$ and $B_{\infty} (t)$ be symmetric
operators in $\bbbr^{2n}$, depending continuously on
$t\in [0,T]$, such that
$A_{\infty} (t) \le B_{\infty} (t)$ for all $t$.

A Borelian function
$H: [0,T]\times \bbbr^{2n} \to \bbbr$
is called
$\left(A_{\infty} ,B_{\infty}\right)$-{\it subquadratic at infinity}
if there exists a function $N(t,x)$ such that:
\begin{equation}
  H (t,x) = \frac{1}{2} \left(A_{\infty} (t) x,x\right) + N(t,x)
\end{equation}
\begin{equation}
  \forall t\ ,\ \ \ N(t,x)\ \ \ \ \
  {\rm is\ convex\ with\  respect\  to}\ \ x
\end{equation}
\begin{equation}
  N(t,x) \ge n\left(\left\|x\right\|\right)\ \ \ \ \
  {\rm with}\ \ n(s)s^{-1}\to +\infty\ \ {\rm as}\ \ s\to +\infty
\end{equation}
\begin{equation}
  \exists c\in \bbbr\ :\ \ \ H (t,x) \le
  \frac{1}{2} \left(B_{\infty} (t) x,x\right) + c\ \ \ \forall x\ .
\end{equation}

If $A_{\infty} (t) = a_{\infty} I$ and
$B_{\infty} (t) = b_{\infty} I$, with
$a_{\infty} \le b_{\infty} \in \bbbr$,
we shall say that $H$ is
$\left(a_{\infty},b_{\infty}\right)$-subquadratic
at infinity. As an example, the function
$\left\|x\right\|^{\alpha}$, with
$1\le \alpha < 2$, is $(0,\varepsilon )$-subquadratic at infinity
for every $\varepsilon > 0$. Similarly, the Hamiltonian
\begin{equation}
H (t,x) = \frac{1}{2} k \left\|k\right\|^{2} +\left\|x\right\|^{\alpha}
\end{equation}
is $(k,k+\varepsilon )$-subquadratic for every $\varepsilon > 0$.
Note that, if $k<0$, it is not convex.
\end{definition}
%

\paragraph{Notes and Comments.}
The first results on subharmonics were
obtained by Rabinowitz in \cite{2fost:kes}, who showed the existence of
infinitely many subharmonics both in the subquadratic and superquadratic
case, with suitable growth conditions on $H'$. Again the duality
approach enabled Clarke and Ekeland in \cite{2may:ehr:stein} to treat the
same problem in the convex-subquadratic case, with growth conditions on
$H$ only.

Recently, Michalek and Tarantello (see Michalek, R., Tarantello, G.
\cite{2fost:kes} and Tarantello, G. \cite{2czaj:fitz}) have obtained lower
bound on the number of subharmonics of period $kT$, based on symmetry
considerations and on pinching estimates, as in Sect.~5.2 of this
article.

%
% ---- Bibliography ----
%
\begin{thebibliography}{}

\bibitem [1982]{2smit:wat}
Smith, T.F., Waterman, M.S.: Identification of common molecular subsequences.
J. Mol. Biol. 147, 195?197 (1981). \url{doi:10.1016/0022-2836(81)90087-5}

\bibitem [1978]{2may:ehr:stein}
May, P., Ehrlich, H.-C., Steinke, T.: ZIB structure prediction pipeline:
composing a complex biological workflow through web services.
In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006.
LNCS, vol. 4128, pp. 1148?1158. Springer, Heidelberg (2006).
\url{doi:10.1007/11823285_121}

\bibitem [1988]{2fost:kes}
May, P., Ehrlich, H.-C., Steinke, T.: ZIB structure prediction pipeline:
composing a complex biological workflow through web services.
In: Nagel, W.E., Walter, W.V., Lehner, W. (eds.) Euro-Par 2006.
LNCS, vol. 4128, pp. 1148?1158. Springer, Heidelberg (2006).
\url{doi:10.1007/11823285_121}

\bibitem [(to appear)]{2czaj:fitz}
Czajkowski, K., Fitzgerald, S., Foster, I., Kesselman, C.: Grid information services
for distributed resource sharing. In: 10th IEEE International Symposium
on High Performance Distributed Computing, pp. 181?184. IEEE Press, New York (2001).
\url{doi: 10.1109/HPDC.2001.945188}

\bibitem [1980]{2fo:kes:nic:tue}
Foster, I., Kesselman, C., Nick, J., Tuecke, S.: The physiology of the grid: an open grid services architecture for distributed systems integration. Technical report, Global Grid
Forum (2002)

\bibitem [(last visited 2013)]{onlyurl}
National Center for Biotechnology Information. \url{http://www.ncbi.nlm.nih.gov}

\end{thebibliography}
\clearpage
%\addtocmark[2]{Author Index} % additional numbered TOC entry
\renewcommand{\indexname}{Author Index}
\printautindex

\end{document}

### 2024-09-03 Articles/2023-HilbertQuantum/Submissions/Quantum Studies/HilbertQuantum.tex
Incorporated Francisco's comments

@@
 This is already a problem: the space of wave-functions for $n$ particles is the Hilbert space $L^2(\mathbb{R}^{3n})$.
-For any value of $n$, these spaces have a countable basis (i.e.
-infinite dimensional and separable) and are therefore all mathematically equivalent.
+For any value of $n$, these infinite-dimensional spaces are separable, meaning that they all have a countable basis, and so are unitarily equivalent.
 That is, the state space for one particle is the same as the state space for, say, a billion particles.
@@
 
-Additionally, Schwartz spaces would be physical only if we are modeling a fixed number of degrees of freedom over Euclidean space.
-We would still need a more general definition that would apply to all cases and would reduce to Schwartz spaces in the above case.
+Additionally, Schwartz spaces can only represent physical systems with a fixed number of degrees of freedom over Euclidean space.
+It is not clear how to generalize the definition to cover all physical cases.
 

### 2024-10-02 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
ClassicalLimit: adding current draft from overleaf

\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{assumptionsofphysics}
\usepackage{tikz-cd}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}
\usepackage{pgfplots}
\def\>{\rangle}
\def\<{\langle}

\title{Classical mechanics as high entropy limit}
\author{Gabriele Carcassi, Manuele Landini, Christine A. Aidala}

\begin{document}
\maketitle

\begin{abstract}
We show that classical mechanics can be recovered as the high entropy limit of quantum mechanics. The idea is that the high entropy hides quantum effects, and mixed states of high enough entropy can be approximated with classical distributions. The mathematical limit $\hbar \to 0$ can be reinterpreted as setting the zero entropy of pure states to $-\infty$, in the same way that non-relativistic mechanics can be recovered mathematically with $c \to \infty$. Physically, these limits are more appropriately defined as $S \gg 0$ and $v \ll c$. Both limits can then be understood as approximations independently of what mechanism allows those approximations to be valid. Consequently, the limit presented is independent of what interpretation is chosen for both quantum states and entropy. 
\end{abstract}

\section{Introduction}

%TODO: Status. Probably two ways: one mathematical (hbar to zero) and multiple mechanism (what happens that makes things loook classical) like decoherence and other things

%TODO: Our aim is to recover the mathematical limit for general physical ideas, in a way that is independent of mechanism and interpretation.

%TODO: give brief overview

\section{High entropy and classical states}

As mentioned in the introduction, we believe that classical mechanics should be recovered from quantum mechanics in a way that makes intuitive sense to all those that routinely work with quantum systems. Therefore, let us start with a few qualitative considerations that hint at the connection between classical mechanics and high entropy.

\subsection{Producing quantum states}

TODO: qui sarebbe bello mettere un sacco di esempi/citazioni

One of the experimental challenges in quantum mechanics is producing states that are ``quantum enough'' to exhibit quantum properties. What we want to show that all these problems can ultimately be understood as reducing the entropy of the initial state.

Coherence is probably one of the direct and most important properties. It has been shown that coherence can be maintained over long distances and among large number of constituents, meaning that quantum systems are not necessarily small or made of few components. However, it is also established how coherence can be quickly lost through interaction with the environment, trough decoherence. Since decoherence does increase the entropy on the system, it represents one mechanism to reach the high entropy limit. This is in line with our result.

In experimental practice, many quantum effects (e.g.~superconductivity, topological insulators, quantum hall effect, ...) are harder or impossible to achieve at high temperature. The thermal noise can in fact break the coherence of the system. Note that entropy is a monotonic function of temperature, meaning that decreasing temperature does mean decreasing entropy of the system. This is in line with our result.

Some quantum effects can be replicated at higher temperature given a high pressure. This is the case, for example, in some superconductive materials. High pressure corresponds to low entropy, which is in line with our result.

To produce Bose-Einstein condensates, one needs a high density in phase space, which means both low entropy (to decrease the position spread) and low temperature (to decrease the momentum spread). This, again, is in line with our result.
{\color{blue} Something of interest: For a trapped condensate, the PSD at the center and S are related by $PSD=e^{5/2+\gamma-S/N}$, where $\gamma$ is the virial coefficient for the trapping potential (Pinske et al Phys.Rev.Lett. 78, 990 (1997)). This has been used experimentally to condense at constant entropy, by changing the shape of the potential (D. M. Stamper-Kurn, Phys. Rev. Lett. 81, 2194 (1998))}

\subsection{Uncertainty from entropy}

Another hint of the link between high entropy and classical mechanics comes from the relationship between entropy and uncertainty.

In both classical and quantum mechanics, Gaussian states maximize entropy at a fixed uncertainty or, equivalently, minimize uncertainty at fixed entropy. This means that, if we fix the entropy $S$, we will have an uncertainty relationship
\begin{equation}
    \Delta x \Delta p \geq \Sigma(S).
\end{equation}
The specific value of $\Sigma$ will depend on the entropy and on whether we are using classical or quantum mechanics, though the relationship will always be saturated by Gaussian states with no correlation between $x$ and $p$.

In classical mechanics, the relationship between entropy and uncertainty for Gaussian states is\cite{Cover_Thomas_2006,Pathria_Beale_2022}
\begin{equation}
    S_C(\Sigma) = \ln \left(2 \pi e \frac{\Sigma}{h}\right) = \ln \left(\frac{\Sigma}{\hbar}\right) + 1.
\end{equation}
In quantum mechanics it is\cite{weedbrook2012gaussian}
\begin{equation}
S_Q(\Sigma) = \left( \frac{\Sigma}{\hbar} + \frac{1}{2} \right) \ln \left( \frac{\Sigma}{\hbar} + \frac{1}{2} \right) - \left( \frac{\Sigma}{\hbar} - \frac{1}{2} \right) \ln \left( \frac{\Sigma}{\hbar} - \frac{1}{2} \right).
\end{equation}

\begin{figure}
    \centering
\begin{tikzpicture}
\begin{axis}[
height=7cm,
width=\linewidth*0.8,
grid=both,
grid style={line width=.1pt, draw=gray!25},
axis lines=middle,
xlabel = \(\Sigma\),
ylabel = \(S\),
legend style={at={(0.95,0.65)},anchor=east},
]
\addplot[blue,samples=150,domain=0.1:8.5] {ln(x)+1};
\addlegendentry{\(\text{classical}\)}
\addplot[red,samples=150,domain=0.5:8.5] {(x+1/2)*ln(x+1/2)-(x-1/2)*ln(x-1/2)};
\addlegendentry{\(\text{quantum}\)}
\addplot[green,samples=150,domain=0.5:8.5] {(ln(x)+1) -((x+1/2)*ln(x+1/2)-(x-1/2)*ln(x-1/2))};
\addlegendentry{\(\text{difference}\)}
\end{axis}
\end{tikzpicture}
    \caption{Entropy $S$ in nats as a function of uncertainty $\Sigma$ in units of $\hbar$. Uncertainty is measured in units of $\hbar$. In blue, the classical case $\ln(\Sigma) + 1$. In red, the quantum case $\left( \Sigma + \frac{1}{2} \right) \ln \left( \Sigma + \frac{1}{2} \right) - \left( \Sigma - \frac{1}{2} \right) \ln \left( \Sigma - \frac{1}{2} \right)$. In green the difference betwen the two.}
    \label{fig:uncertainty}
\end{figure}

In figure \ref{fig:uncertainty} we can see that the classical and quantum cases are very close when the uncertainty is even just a few units of $\hbar$. Things diverge at about two units of $\hbar$: in quantum mechanics the entropy decreases faster and reaches zero at $\hbar/2$, the bound for the Heisenberg uncertainty principle; classical mechanics reaches zero entropy for $\hbar/e$, and then continues in the negative region.

Note that negative entropy is problematic. It directly contradicts the third law of thermodynamics, but we can see other, less axiomatic, problems. Given that entropy is additive under composition of independent systems, in the negative region the entropy decreases as the number of particles increases. This means that, at low temperature, zero particles give the highest entropy. This is one problem with classical theory that is corrected by quantum mechanics.

Studying the relationship between uncertainty and entropy, then, reinforces the idea that quantum mechanics is required at low entropy much like relativity is required at high speeds.

\subsection{Entropic aliasing}

Note that, in our approach, the mechanism that increases the entropy is irrelevant: any will do. The intrinsic geometry of quantum mixed states makes uncertainty over classical or quantum properties look the same. It is this aliasing of different preparations that drives the classical limit, as we can see even in a standard two slit experiment setup.

In the simplest case, we can imagine the particle either passing through the left or right slit, which corresponds to the two pure states $|L\>$ and $|R\>$. These identify a two-state system, a qubit. We can now imagine equal superpositions of the two states, $\frac{1}{2}|L\> + \frac{1}{2} e^{-\imath \theta} |R\>$ where $\theta$ is the difference in phase between the two components. If $\theta$ is zero, we have the state $|+\> = \frac{1}{2}|L\> + \frac{1}{2} |R\>$ and the resulting interference pattern will have a peak in the middle of the screen. If $\theta$ is $\pi$, we have the state $|-\> = \frac{1}{2}|L\> + e^{-\imath \pi}\frac{1}{2} |R\>$ and the resulting interference pattern will have a valley in the middle of the screen.

As expected, $|+\>$ cannot be understood as a probability distribution, as a mixture, over $|L\>$ and $|R\>$. The interference pattern, the quantum feature, cannot be understood as a classical distribution over a path, a classical feature. However, an equal mixture of $|+\>$ and $|-\>$ is the maximally mixed state, which is equal to an equal mixture of $|L\>$ and $|R\>$. That is, the case where we randomize the phase is indistinguishable from the case where we randomize the path. In other words, uncertainty over a quantum feature behaves like uncertainty over a classical feature, regardless of the source of the uncertainty.

In the Bloch ball, states at equal entropy form a series of concentric spheres. As we increase the entropy, the points of these spheres become closer and closer, and therefore the error in using a mixture of $|L\>$ and $|R\>$ instead of the actual state becomes smaller. The same happens for the space of mixtures of any quantum system.

Mathematically, the trace distance $T(\rho,\sigma)$ between two mixed states $\rho$ and $\sigma$ will decrease under a completely positive trace-preserving (CPTP) map $M$ that increases the entropy of all states. That is, $T(M(\rho),M(\sigma)) < T(\rho,\sigma)$.\footnote{TODO Man: citation needed} If we call $\mathcal{G}$ the space of all mixtures of Gaussian states, given an arbitrary mixed state $\rho$, we will find a Gaussian state $\sigma_{\rho} \in \mathcal{G}$ that is closest to $\rho$. As the entropy of $\rho$ increases, $\sigma_{\rho}$ will be get closer to $\rho$.

What we described here is just a geometric property of the space of quantum mixed states, and it is independent on the source of entropy.

\section{Reinterpreting traditional approaches}

As mentioned, one of advantages of our approach is that it is compatible with more traditional approaches, including taking the limit $\hbar \to 0$. What one finds is that the math can typically be reinterpreted by taking an equivalent more physically meaningful limit.

For example, following the original paper from Wigner \cite{WignerLimit}, we can consider the Wigner distribution of a system in thermal equilibrium at inverse temperature $\beta$:\footnote{TODO Man: there seem to be some difference from the original paper}
\begin{equation}
    W(x,p)=\int dy e^{i(x+y)p/\hbar}\langle x+y |e^{-\beta\hat{H}}|x-y \rangle e^{-i(x-y)p/\hbar}
\end{equation}
Thermal equilibrium is described by $\hat{\rho}=e^{-\beta \hat{H}}$, which is the mixed state that maximizes entropy at a given average energy. The entropy of this state is directly connected to the value of $\beta$ such that when $S$ tends to $\infty$, $\beta$ tends to 0.
From this, Wigner considers the transformed Hamiltonian
\begin{equation}
    \tilde{H}=e^{i x p/\hbar}\hat{H} e^{-i x p/\hbar}=\frac{(p+i\hbar\partial/\partial x)^2}{2m}+V(x)=\epsilon(x,p)+i\frac{\hbar p}{m}\frac{\partial}{\partial x}-\frac{\hbar^2}{2m}\frac{\partial^2}{\partial x^2}
\end{equation}
where $\epsilon(x,p)$ is the classical Hamiltonian. The extra term contains the quantum corrections; we will refer to it as $Q(\partial/\partial x)$.\footnote{TODO Man: Q vs Q hat} The Wigner function becomes
\begin{equation}
    W(x,p)=\int dy \langle x+y |e^{-\beta\tilde{H}}|x-y \rangle 
\end{equation}
At this point, Wigner expands this expression in powers of $\hbar$, showing that quantum corrections are at second order. What we can do is instead expand in powers of $\beta$. We can show that quantum corrections are also only found at second order in $\beta$, justifying the classical limit. Let's start by considering the first order expansion, $e^{-\beta\tilde{H}}\simeq 1-\beta(\epsilon+\hat{Q})$. So at first order
we get
\begin{equation}
    W(x,p)\simeq 1-\beta \epsilon(x,p)-\beta\int dy\langle x+y |\hat{Q}|x-y \rangle.
\end{equation}
Focusing on the last term, we can insert a completeness (?) in the momentum eigenbase
\begin{equation}
  \int dy\langle x+y |\hat{Q}|x-y \rangle=\int\int dkdy e^{i2ky} \left(-\frac{\hbar p}{m}k+\frac{\hbar^2}{2m}k^2\right)=\int dk \delta(2k) \left(-\frac{\hbar p}{m}k+\frac{\hbar^2}{2m}k^2\right)=0.
\end{equation}
The first order quantum correction vanishes, showing that the classical approximation is correct to first order in $\beta$, which corresponds to the limit of large entropy. We can consider the second order correction
\begin{equation}
     W(x,p)\simeq 1-\beta \epsilon(x,p)-\beta^2\epsilon^2(x,p)-\beta^2\int dy\langle x+y |\hat{Q}V(\hat{x})+V(\hat{x}) \hat{Q}|x-y \rangle,
\end{equation}
giving the first nonzero quantum correction to the Wigner function.

\section{The high entropy limit}

We are now going to develop the limit in its most general formulation. TODO reword. Since the maximum entropy attainable by a quantum system is the logarithm of the dimension of the corresponding Hilbert space, the high entropy limit exists only for infinite-dimensional spaces.\footnote{TODO Man: add reference} This explains why spin, which lives in a finite-dimensional space, is an intrinsically quantum property. Therefore we are going to concentrate on the case of a single degree of freedom identified by position and momentum. The case of multiple degrees of freedom can be recovered by increasing the entropy of all DOFs at the same time. That is, increasing the entropy of each DOF independently, without introducing correlations.

The overall argument can be broken down into the following steps:
\begin{enumerate}
    \item characterize entropy-increasing maps in classical mechanics for a single DOF, noting that they can all be understood as Hamiltonian evolutions followed by a pure stretching map, one that stretches position and momentum by the same factor $\sqrt{\lambda}$
    \item show that a pure stretching map in quantum mechanics cannot be defined over symmetrized operator averages or normal ordering
    \item show that it can be defined over anti-normal ordering; show that if we rescale the zero of entropy, the commutators become $[X, P]= \frac{\imath \hbar}{\lambda}$, which makes the limit $\lambda \to \infty$ mathematically equivalent to $\hbar \to 0$
    \item show how, in the limit, the Wigner $W$ and Husimi $Q$ distributions become closer and closer
\end{enumerate}

\subsection{Stretching classical phase space}

Before looking at the quantum case, let us study the high entropy limit of classical mechanics. We are looking for all those transformations that increase the entropy of all states by the same amount.

We are going to study the one-dimensional case, therefore let us call $\mathcal{M} = (\mathbb{R}^2, \omega)$ the phase space for a single degree of freedom, where $\omega$ is the associated symplectic form.\footnote{The symplectic form allows us to define the count of states in infinitesimal regions, and therefore defines the entropy as well.} Suppose we have a map $R : \mathcal{M} \to \mathcal{M}$ that acts on phase space. This map will also act on probability distributions defined on $\mathcal{M}$ by moving them point to point. We require that $R$ increases the entropy of each distribution $\rho$ by a set value $\Delta S = \log \lambda > 0$. That is
\begin{equation}
S(R(\rho)) = S(\rho) + \log \lambda,
\end{equation}
where $S(\rho) = - \int_{\mathcal{M}} \rho \log \rho dx dp$ and $R(\rho)$ is the distribution as it is transformed through the map.

For a generic transformation, the increase of entropy is given by the expectation of the Jacobian determinant $\Delta S =\int_M \rho \log |\partial_a R^b| dxdp$. Since the increase has to be the same for all distributions, we must have $|\partial_a R^b| = \lambda > 1$.\footnote{The case of a negative Jacobian is a reflection in phase space, which cannot be achieved through a continuous evolution in time and is therefore discarded.} Recall that the Jacobian determinant tells us how an infinitesimal volume scales through the transformation, and therefore a map that increases entropy uniformly is exactly a map that stretches phase space uniformly. We call such a map a \textbf{stretching map}.

Suppose, in fact, that $\rho$ is a uniform distribution over a region of area $W_1$. Then we have $S(\rho)= \log W_1$. Applying the stretching map, $\rho$ will transform into a uniform distribution over a region of area $W_{\lambda} = \lambda W_1$. The final entropy is therefore $S(R(\rho)) = \log \lambda W_1 = \log W_1 + \log \lambda$. The factor $\lambda$, then, can be understood either as the ratio between initial and final areas, or as the exponential of the entropy increase. Therefore studying the high entropy limit means studying what happens under stretching maps in the limit $\lambda \to \infty$.

Note that the transformation is not a canonical transformation. Canonical transformations, those that can be generated by Hamiltonian evolution, preserve areas in phase space and conserve entropy. In fact, for a single degree of freedom, canonical transformations and volume-preserving maps coincide. This allows us to show that all stretching maps can be written as $R = T \circ U$, where $U$ is a canonical transformation and $T$ is a \textbf{pure stretching map} defined as
\begin{equation}
    T(x,p) \mapsto (\sqrt{\lambda} x, \sqrt{\lambda} p)
\end{equation}
where $\lambda = (1,\infty)$. That is, any stretching map can be understood as first performing a canonical transformation that preserves entropy followed by a pure stretch of position and momentum.

To see how this works, note that given a stretching map $R$, we can always write $U = T^{-1} \circ R$. Since $R$ stretches phase space everywhere by a factor $\lambda$ and $T^{-1}$ shrinks it by the same factor, $U$ preserves areas and is a canonical map, since we restricted ourselves to the one-dimensional case.\footnote{To extend to the general case, the stretching map must not only stretch the total volume, but areas in each DOF. Mathematically, this means rescaling the symplectic form $\omega$.} Therefore $R = T \circ U$, which means we only need to study $T$ to characterize all maps that increase entropy uniformly.

We can alternatively characterize stretching maps by how the  Poisson brackets transform. Note that
\begin{equation}
    \{R(x),R(p)\} = \partial_x R(x) \partial_p R(p) - \partial_x R(p) \partial_p R(x) = | \partial_a R^b | = \lambda.
\end{equation}
That is, the Poisson bracket of the transformed position and momentum is the Jacobian determinant of the transformation, which is $\lambda$. Since this is an equality, all maps that satisfy the above transformation of the Poisson brackets are stretching maps.

For a pure stretching map, we can provide a more specific characterization. Note, in fact, that a distribution with compact support is always fully characterized by all its central moments, that is the expectations for all polynomials of position and momentum. Therefore, once we know how the central moments transform through the map, we know how all distributions with compact support transform and therefore we fully characterize the map. This means that $T$ is a pure stretching map if and only if
\begin{equation}
    \langle T(x^np^m) \rangle = (\sqrt{\lambda})^{(n+m)} \langle x^n p^m \rangle.
\end{equation}

\subsection{Operator ordering}

Ideally, we would want to treat the quantum case similarly to the classical case. We would look for those quantum channels (i.e.~CPTP maps) $R$ that increase the entropy of all mixed states by the same amount, we would show that they can be decomposed into a unitary part and a pure stretching part and so on. In practice, the quantum version of the problem is much harder. First of all, a map that increases the entropy of all quantum states cannot be invertible: pure states have the lowest entropy and therefore no state of lower entropy can be mapped to them. Secondly, since our map increases entropy, it does not in general preserve products. In fact, if we had $R(AB) = R(A)R(B)$, then $R([A,B]) = [R(A),R(B)]$, which would mean $R$ is unitary and therefore preserves entropy. Since, in general, $R(AB) \neq R(A)R(B)$, different permutations of polynomials will be transformed differently: the operator ordering matters.
%GC: I think this is a simpler, more general argument. It also puts more in focus why I kept getting confused from the other argument

% Note https://link.springer.com/article/10.1007/BF00419590 says entropy does not decrease if L L* \leq L* L. There also seems to be a bound

In this light, we are looking for a map that rescales the expectations of the polynomials for a particular ordering. There are three commonly used orderings: symmetrized averages, normal ordering and anti-normal ordering. The first takes the product of $n$ operators by averaging all possible permutations; the second is in terms of polynomials of the form $(a^\dagger)^n a^m$, where $a=\sqrt{\frac{m\omega}{2\hbar}}(X+\frac{i}{m\omega}P)$ and $a^\dagger$ are the ladder operators; the last is in terms of $a^n (a^\dagger)^m$. The three orderings are associated, respectively, with a quasi-probability distribution: the Wigner function $W$, the Glauber-Sudarshan $P$ distribution and the Husimi $Q$ distribution. In each case, the expectation under the quasi-probability distribution of polynomials of classical variables returns the expectation of the respective ordering.

Interestingly, not all orderings will allow a pure stretching map $T$ that rescales all expectations and increases the entropy of all states. Let us consider the symmetrized average case. We are looking for a map $T_W$ for which
\begin{equation}
\langle T_W(\Pi(\underbrace{X, ..., X}_{\text{n times}}, \underbrace{P, ..., P}_{\text{m times}})) \rangle = (\sqrt{\lambda})^{(n+m)} \langle \Pi(\underbrace{X, ..., X}_{\text{n times}}, \underbrace{P, ..., P}_{\text{m times}})\rangle.
\end{equation}
where $\Pi(A_1, A_2, \ldots, A_n)  = \frac{1}{n!} \sum_{\pi}  A_{\pi(1)} A_{\pi(2)} \cdots A_{\pi(n)}$ is the average of the products for each permutation $\pi$. These averages correspond to the expectation calculated through the Wigner function $W(x,p)$
\begin{equation}
     \langle \Pi(\underbrace{X, ..., X}_{\text{n times}}, \underbrace{P, ..., P}_{\text{m times}})\rangle = \int_M x^n p^m W(x, p) dx dp.
\end{equation}
The map $T_W$, then, would correspond to a pure stretching map on the Wigner function. However, this cannot work. Wigner functions can have regions with negative values, but the size of these regions cannot exceed a few units of $\hbar$.\cite{kenfack2004negativity} The size of these negative regions would increase under $T_W$, giving functions that do not correspond to a quantum state. Therefore we cannot find a pure stretching map in the symmetrized average operator ordering.

Let us now consider a pure stretching map $T_P$ in normal ordering. This would have:
\begin{equation}
\langle T_P((a^\dagger)^n a^m) \rangle = (\sqrt{\lambda})^{(n+m)} \langle (a^\dagger)^n a^m \rangle.
\end{equation}
For the vacuum state we have $a|0\rangle=0$, which means the mean value of all observables in normal ordering for the vacuum is zero. These would remain unchanged by $T_P$. The vacuum state, then, would not change and therefore the map would not increase entropy for all states. The normal ordering is ruled out as well. {\color{blue} thinking of it this way, you can also use the same argument as before. The P distribution also has negative regions that would stretch.} {\color{red} do we know they must have a maximum size? I haven't seen anything about it.}

We turn our attention to the anti-normal ordering and a map $T_Q$ such that:
\begin{equation}
\langle T_Q(a^n(a^\dagger)^m) \rangle = (\sqrt{\lambda})^{(n+m)} \langle a^n(a^\dagger)^m \rangle.
\end{equation}
This ordering solves the previous problem of the vacuum. The anti-normal ordering is connected to the Husimi $Q$ distribution by
\begin{equation}
    \langle a^n (a^\dagger)^m\rangle=\int \alpha^n(\alpha^*)^m Q_1(\alpha)d^2\alpha.
\end{equation}
The map $T_Q$, then, corresponds to a pure stretching map on the space where $Q$ is defined. That is,
\begin{equation}
T_Q(Q_1(\alpha)) \equiv Q_\lambda(\alpha) = \frac{1}{\lambda}Q_1\left(\frac{\alpha}{\sqrt{\lambda}}\right).
\end{equation}
We indicate $Q_1$ as the initial unstretched distribution and $Q_\lambda$ the final stretched distribution by a factor of $\lambda$. We can in fact verify that 
\begin{equation}
\int \alpha^n(\alpha^*)^m Q_\lambda(\alpha)d^2\alpha=\int \alpha^n(\alpha^*)^m \frac{1}{\lambda}Q_1\left(\frac{\alpha}{\sqrt{\lambda}}\right)d^2\alpha=\int \sqrt{\lambda}^{n+m}\beta^n(\beta^*)^m Q_1(\beta)d^2\beta.
\end{equation}
Since the Husimi distribution is non-negative, this avoids the issue presented by the Wigner function.

The anti-normal ordering, then, is a potential candidate. We now need to show that a pure stretching map $T_Q$ actually exists.

\subsection{Pure quantum stretching map}

If pure quantum stretching maps $T_Q$ exist, they must be CPTP maps as they must transform mixed states into mixed states. Moreover, since they have to increase entropy for all states, they must be describing an open quantum system. We thus show that pure stretching maps can be expressed as the solution of a master equation in Lindblad form and the Heisenberg picture. We start with
\begin{equation}
	\frac{d}{d t} X=\frac{i}{\hbar}[H,X]+\sum_i \gamma_i \left(L_i^\dagger X L_i-\frac{1}{2}\left\{L_i^\dagger L_i,X\right\}\right)
\end{equation}
where X is an operator, $L_i$ are the jump operators and $\gamma_i$ are positive parameters. To define $T_Q$ for a single DOF we set $\gamma_i=\{\gamma\}$ and $L_i=\{a^\dagger\}$. We also set $H=0$ since $T_Q$ should be purely dissipative. The equation of motion of $a$ reduces to
\begin{equation}
	\frac{d}{d t} a=\frac{\gamma}{2} a
\end{equation}
with solution $a(t)=e^{\frac{\gamma}{2} t} a = \sqrt{\lambda} a$, where we make the identification $\lambda=e^{\gamma t}$. 
In general, we find
\begin{equation}
	\frac{d}{d t} a^n (a^{\dagger})^m=\frac{\gamma}{2}(n+m)a^n (a^{\dagger})^m
\end{equation}
meaning that this evolution realizes the stretching map in anti-normal ordering. That is, $(a^n (a^{\dagger})^m)(t)=e^{\frac{\gamma}{2} t (n+m)} a^n (a^{\dagger})^m = (\sqrt{\lambda})^{(n+m)} a^n (a^{\dagger})^m$.\footnote{Note that for the choice of $L_i=\{a\}$, we get the opposite behavior, characterized by
\begin{equation}
	\frac{d}{d t} (a^{\dagger})^n a^m=-\frac{\gamma}{2}(n+m)(a^{\dagger})^n a^m
\end{equation}
leading to shrinking ($\lambda\leq1$) of all observables taken in normal ordering.}

This shows that a map $T_Q$ that stretches the $Q$ distribution can be understood as a purely dissipative process that runs for a time $\Delta t$ with $\gamma_i = \left\{ \frac{\log \lambda} {\Delta t} \right\}$ and $L_i = \{ a^\dagger\}$, and therefore it is a CPTP map. Note that, for the same DOF, $a^\dagger$ is not uniquely fixed as it depends on the parameter $m\omega$. Moreover, ladder operators with different values of the parameter $m\omega$ are not going to commute, therefore will lead to different transformations. That is, while we fixed the ordering of the operators, the non-commutative nature of quantum observables still implies a choice within a one-parameter family of operators.  However, note that a linear transformation that stretches $X$ and shrinks $P$
\begin{equation}
	U(X, P) = \left(\sqrt{\alpha} X, \frac{1}{\sqrt{\alpha}} P\right)
\end{equation}
is a unitary transformation. Under this map, $U(a) = \sqrt{\frac{m\omega\alpha}{2\hbar}}(X+\frac{i}{m\omega\alpha}P)$, which means that $m\omega$ can be changed through a unitary operator. Consistently with the classical definitions, we can define $R_Q = T_Q \circ U$ to be a quantum stretching map. Therefore, even if fixing the operator ordering does not pick a unique map, the behavior when entropy increases is the same.

Lastly, we need to show that $T_Q$ increases entropy for all states. Note that a Lindblad operator increases entropy for all states if\cite{Benatti1988}
\begin{equation}
	\sum_i L_i.L_i^\dagger \leq \sum_i L_i^\dagger L_i.
\end{equation}
In our case, this reduces to 
\begin{equation}
	a^\dagger a \leq a a^\dagger.
\end{equation}
Since $[a, a^\dagger] = a a^\dagger - a^\dagger a = I \geq 0$, our map increases entropy for all states.\footnote{Note that for the choice $L_i=\{a\}$ the inequality is not satisfied.}

We can now look at the effect of the map on $X$, $P$ and their commutator. We have
\begin{align}
    T_Q(a) &= \sqrt{\lambda} a \\ 
    T_Q(a^\dagger) &= \sqrt{\lambda} a^\dagger \\ 
    T_Q(X) &= T_Q\left(\sqrt{\frac{\hbar}{2m\omega}}(a^\dagger + a)\right) = \sqrt{\lambda} X \\
    T_Q(P) &= T_Q(\imath\sqrt{\frac{\hbar m \omega}{2}}(a^\dagger - a)) = \sqrt{\lambda} P \\
    [T_Q(X), T_Q(P)] &= \lambda [X, P].
\end{align}
In other words, the effect of $T_Q$ is to increase the commutator between $X$ and $P$.

We can study the limit in two ways. The most straightforward would be see how the space changes as the commutators $[T_Q(X), T_Q(P)]= \lambda \imath \hbar$ go to infinity. This corresponds to keeping the entropy of the pure states fixed at zero and studying the space of states at high entropy. Alternatively, we can see how the space itself changes from the perspective of the state that is increasing in entropy. That is, we rescale the entropy of pure states to lower and lower values, while keeping the entropy of the mixed state fixed. To do that, we keep $[T_Q(X), T_Q(P)] = \imath \hbar$ constant and we study the limit as $[X, P]= \frac{\imath \hbar}{\lambda}$ goes to zero. Formally, this is equivalent to taking the limit $\hbar \to 0$, which is the group contraction that morphs the Moyal bracket Lie algebra to the Poisson bracket Lie algebra.\cite{saletan1961contraction, inonu1953contraction} That is, as $\lambda$ increases, the states will see the structure of quantum mechanics becoming closer and closer to the structure of classical mechanics.

%TODO: still missing a group contraction reference for \hbar \to 0

Note that the above argument works for any map $T_Q \circ U$ that combines the pure stretching map with an arbitrary unitary evolution $U$. Also, while we have not shown that anti-normal ordering is the only ordering that allows a stretching map, it will need to have the same effect on $X$ and $P$, leading to the same limit. The question of whether all maps that perform the same group contraction can expressed as $T_Q \circ U$ remains open. However, they will factorize in that fashion in the limit, since the classical limit can always be factorized as pure stretching map following a symplectomorphism. Similarly, the question of whether $T_Q$ increases the entropy of all states by the same amount remains open. However, it will do so in the limit since $T_Q$ will become closer and closer to a pure classical stretching map. Lastly, we have not shown that all transformations that increase entropy must stretch phase space. Note, however, that a finite region of phase space can only hold mixed states with finite entropy. Therefore a map that does not stretch some region of phase space will necessarily lead to states whose entropy will not go to infinity as the map is reapplied over and over. Regardless, our ultimate goal is not to study all possible ways entropy can be increased. Our goal is simply to show that, as entropy increases, the classical description becomes a suitable approximation.

To sum up, we have found a pure quantum stretching map $T_Q$. It is a CPTP map. It rescales operators in anti-normal ordering. It increases entropy for all states. It recovers classical mechanics in the limit. This means that the space of quantum states, as entropy increases, becomes more and more classical. All of this is done under general conditions, and is independent of the mechanism that performs the entropy increase.

%Since the canonical operators x and p are simply linear combinations of $a$ and $a^\dagger$, their mean values will stretch as required, with stretching parameter given by $\lambda=e^{\gamma t}$. It is not so hard to show that the scaling applies to all operator combinations in anti-normal ordering. 
%If one is careful, it's clear that this cannot work forever. We can stretch the first moment of the operators: $a_i^{(\dagger)}(t)=\sqrt{\lambda} a_i^{(\dagger)}(0)$. But already for the second moment this breaks. We find: $(a^\dagger a)(t)=\lambda((a^\dagger a)(0)+1)-1$. The exact stretching is only recovered in the limit of  $\lambda\rightarrow\infty$. On the other hand, if we consider the anti-normal ordering $(a a^\dagger)(t)=\lambda(a a^\dagger)(0)$. This means that the map we have satisfies stretching in anti-normal ordering.

%The correct distribution function to use under anti-normal ordering is the Husimi Q-distribution. This makes perfect sense. This is the only one that is positive-defined. Therefore it's the only one that can be stretched without issues due to negative regions

%TODO: show that such a map exist and it increases the entropy by a fixed amount (DONE?)

% Possible resource https://link.springer.com/article/10.1007/BF00419590 to show map increase entropy

\subsection{Stretching the Wigner distribution}

To understand what happens physically during the group contraction, let us see how the Wigner W distribution changes under a pure quantum stretching map. To do that, we can use the fact that Q is the Weierstrass transform of $W$. That is:
\begin{equation}
    Q_\lambda(\alpha)=\frac{2}{\pi}\int W_\lambda(\beta)e^{-2|\alpha-\beta|^2}d^2\beta.
\end{equation}
Considering the evolution of $Q$, we get
\begin{equation}
    \int W_\lambda(\beta)e^{-2|\alpha-\beta|^2}d^2\beta=\frac{1}{\lambda}\int W_1(\beta)e^{-2|\alpha/\sqrt{\lambda}-\beta|^2}d^2\beta.
\end{equation}
We notice that these two equations are scaled convolutions between the Wigner distribution and Gaussian functions. Symbolically:
\begin{equation}
    (W_\lambda \ast G)(\alpha)=\frac{1}{\lambda}(W_1\ast G)\left(\frac{\alpha}{\sqrt{\lambda}}\right)
\end{equation}
where $G(\alpha)=(2/\pi) e^{-2|\alpha|^2}$ indicates the Gaussian function. To get rid of the convolution, we now take a Fourier transform of both sides. Keeping in mind that the Fourier transform of a general function $f$ behaves under scaling according to
\begin{equation}
    F\left(\frac{1}{\lambda}f\left(\frac{\alpha}{\sqrt{\lambda}}\right)\right)(k)=F(f(\alpha))(\sqrt{\lambda}k)
\end{equation} and using the convolution theorem, we get (TODO Man: check)
\begin{equation}
  F(W_\lambda)(k)F(G)(k)=F\left(\frac{1}{\lambda}W_1\left(\frac{\beta}{\sqrt{\lambda}}\right)\right)(k)F(G)(\sqrt{\lambda}k).  
\end{equation}
We now notice that (TODO Man: add (k) in the end?)
\begin{equation}
    \frac{F(G)(\sqrt{\lambda}k)}{F(G)(k)}=\frac{e^{-\lambda|k|^2/8}}{e^{-|k|^2/8}}=e^{-(\lambda-1)|k|^2/8}=F(G)(\sqrt{\lambda-1}k)=F\left(\frac{1}{\lambda-1}G\left(\frac{\beta}{\sqrt{\lambda-1}}\right)\right),
\end{equation}
finally giving
\begin{equation}
    W_\lambda(\beta)=F^{-1}\left(F\left(\frac{1}{\lambda}W_1\left(\frac{\beta}{\sqrt{\lambda}}\right)\right)F\left(\frac{1}{\lambda-1}G\left(\frac{\beta}{\sqrt{\lambda-1}}\right)\right)\right)=\frac{1}{\lambda(\lambda-1)}W_1\left(\frac{\beta}{\sqrt{\lambda}}\right) \ast G\left(\frac{\beta}{\sqrt{\lambda-1}}\right)
    \label{Fourier}
\end{equation}
Writing this down explicitly
\begin{equation}
   W_\lambda(\beta)=\frac{2}{\pi\lambda(\lambda-1)}\int W_1\left(\frac{\alpha}{\sqrt{\lambda}}\right)e^{-\frac{2}{\lambda-1}\left|\alpha-\beta\right|^2}d^2\alpha.
\end{equation}
A similar calculation gives the Glauber-Sudarshan $P$ distribution, as this one is also obtained from a Weierstrass transfrom of $Q$.

It is interesting to consider what happens to the negative regions of $W$ under the stretching map. We know that $W$ can have negative regions, but their size is limited by the uncertainty principle. In fact, convolving $W$ with a 2D Gaussian with unitary spread, as in the definition of $Q$, returns a function that is never negative. In the limit $\lambda\gg1$, the formula for $W_\lambda$ reduces to
\begin{equation}
     W_\lambda(\beta)\rightarrow_{\lambda\gg1}\frac{2}{\pi\lambda^2}\int W_1\left(\frac{\alpha}{\sqrt{\lambda}}\right)e^{-\frac{2}{\lambda}\left|\alpha-\beta\right|^2}d^2\alpha=Q_\lambda(\beta).
\end{equation}
Therefore, while negative regions can be in principle found at any finite $\lambda$, $W$ tends to a positive function in the limit. As usual for the $W$ distribution, the phase space size of negative regions is limited to $\hbar$ by the uncertainty principle. The weight of the function in such regions is limited between $\pm 2/\hbar$ for pure states. The effect of the stretching map is to reduce this bound to $\pm 2/(\lambda\hbar)$ for large values of $\lambda$. A clear interpretation can be made by working directly on the Fourier transform of $W$; see Eq.~(\ref{Fourier}). The function $F(W_\lambda)$ is a scaled version of $F(W_1)$, with a Gaussian filter applied to it. Quantum information in $W_1$ is carried by spectral weights with k-vectors larger than 1. The bandwidth of the Gaussian filter is given by $\lambda/(\lambda-1)$. This cutoff approaches 1 in the limit, filtering away the interference terms.

We have thus shown that, in the limit of high entropy, the Wigner distribution can be approximated by a positive function. We now need to show that the evolution can be approximated by classical Hamiltonian evolution. An intuitive way to understand why this works is to look at the evolution of the Wigner function under a Hamiltonian $H = p^2/2m + V$ where the potentials $V$ are analytic. We have~\cite{hillery1984distribution}
\begin{equation}
\begin{aligned}
    \frac{d}{dt} W &= \{\{H, W\}\} \\
    &= \{H, W\} + \sum_n \frac{\hbar^{2n} (-1)^n}{(2n+1)! \, 2^{2n}} \partial_x^{(2n+1)} V \partial_p^{(2n+1)} W
\end{aligned}
\end{equation}
The evolution is in terms of the Moyal bracket $\{\{H,W\}\}$ which can be expanded in orders of $\hbar$ with the Poisson bracket $\{H,W\}$ as the leading term. In our limit, $\hbar$ is constant but as the function stretches all derivatives of $W$ decrease. Therefore the first term becomes dominant.

\section{Conclusion}

We have seen that classical mechanics can be recovered as the high entropy limit of quantum mechanics. That is, states of high entropy are better and better approximated by classical distributions over phase space. This approach to the classical limit is independent of mechanism and interpretation, as it does not matter how the entropy is increased or what one believes quantum states to represent: as long as the description is in terms of mixed states of sufficiently high entropy, classical mechanics applies. The approach fits naturally with experimental considerations and other approaches, such as decoherence, and recovers the established mathematical recipe, which is the group contraction for $\hbar \to 0$. Physically, this can be understood as taking the entropy of pure states to minus infinity, which is equivalent to saying that the relative entropy of mixed states goes to plus infinity. This gives a reasonable and precise account of the classical limit: in the same way that non-relativistic mechanics applies to low speed, classical mechanics applies to high entropy.

The limit also gives us additional insights. As we saw, classical states for which $\Delta x \Delta p < \frac{\hbar}{e}$ are states with negative entropy, which imply a breakdown of thermodynamics. It is no wonder, then, electrons cannot fall on the nucleus or that classical mechanics gives us the wrong spectra for black-body radiation. Classical mechanics fails at low entropy. The use of the correspondence principle as guidance for the development of quantum mechanics, then, can be understood as requiring that the new quantized theory reproduces classical mechanics at high entropy.

In this light, we wonder whether it can be proven that quantum mechanics is the only way to fix the low entropy range of classical mechanics. That is, is quantum mechanics the only theory that can recover classical mechanics at high entropy? To us, this is a question whose answer would be interesting regardless of the outcome.

\bibliographystyle{plain}
\bibliography{bibliography}

\end{document}

### 2024-10-09 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: skelton and first part

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}

\usepackage{assumptionsofphysics}
\usepackage{tikz}
\usepackage{breakurl}

\newcommand\hull{\mathrm{hull}}
\newcommand\stcap{\mathrm{scap}}
\newcommand\fraction{\mathrm{frac}}
\newcommand\frcap{\mathrm{fcap}}

\newcommand{\ens}[1][e] {\mathsf{#1}} % Ensemble
\newcommand{\Ens}[1][E] {\mathcal{#1}} % Ensemble space

\def\>{\rangle}
\def\<{\langle}
\DeclareMathOperator{\erf}{erf}

\begin{document}

\title{A non-additive generalization of probability theory \\for quantum mechanics and beyond}
\author{Gabriele Carcassi}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\author{Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
	We present a physically motivated generalization of probability theory that is suitable for classical mechanics, quantum mechanics and any future physical theory that allows a statistical description. The goal is to put the use of classical and quantum probability in a broader context, and show how the current mathematical structures are likely not suitable to solve the open problems in the foundations of physics. For the more mathematically inclined, we will point to areas where new math or generalization of established math are needed. For the more philosophically incline, we will point to areas where further conceptual work is needed.

	%Given a generic space of ensembles, we can define the fraction capacity as the maximum fraction of a particular ensemble that can be understood as a mixture of ensembles from a given set. This gives a non-additive (i.e. fuzzy) measure that reduces to a probability (i.e. additive) measure for classical spaces and for quantum measurements. We can also define the state capacity as the exponential of the maximum entropy reachable by a mixture of ensembles from a given set. This is also a non-additive measure and reduces to the Liouville measure in classical mechanics and the Hilbert space dimensions for subspaces of quantum mechanics. Conceptually, it gives us a notion of probability that is both theory and interpretation independent. Mathematically, it gives us a measure theoretic generalization of probability. Physically, it may allow to find new way to understand current theory and tool to investigate new ones. The purpose of this paper is to show the core ideas and present questions that may be developed on the mathematical, physical and philosophical side.
\end{abstract}

\maketitle

\section{Introduction}
Paper for physicists:
* argue that we need to develop mathematical structures from physical requirements
* create a theory of probability that works for classical, quantum mechanics and beyond
* inverted structure

\section{Ensemble spaces}

We are looking to find the minimal set of axioms required by a physical theory that allows statistical descriptions. Such a theory must, at the very least, provide us with a notion of ensemble, and a space of ensembles that are allowed in the theory. An ensemble can be understood as the infinite collection of the outputs of a particular preparation procedure. The ensemble space of classical mechanics, for example, is the space of continuous probability distributions of phase space. In quantum mechanics, instead, it is the space of density operators. While the ensemble space for each theory will be in principle different, some basic features will remain the same. We are looking for these basic features, necessary elements that any ensemble space must have.

Note that there may good philosophical grounds to require that a physical theory allows for statistical descriptions. If a physical theory has to be testable, and repeatedly so, then we fail to see how it could be so without providing statistical tools. Every state preparation and measurement, in fact, is typically statistical in nature. The development of such arguments would be interesting to us.

We have three basic requirements on ensembles: first, they have to be identifiable experimentally; second, they need to allow mixtures; lastly, they need an entropy well defined. Let us go through these items one by one.

\subsection{Experimental verifiability}

\subsection{Statistical mixtures}
Another basic requirement of an ensemble space is ability to prepare a \textbf{mixture} of two ensembles. If $\ens[a], \ens[b] \in \Ens$ are two ensembles, and $p \in [0,1]$ a weight, we will find the $\ens = p \ens[a] + (1-p) \ens[b] \in \Ens$ that describes a process that selects the first ensemble over the second $p$ percent of the times. Mathematically, the ensemble space is endowed with a \textbf{convex structure}.

Ultimately, the convex structure is responsible for all linear structures we have in physics. Both in classical and quantum mechanics, the ensemble space is not just a convex set, but a convex subset of a vector space. In those cases, in fact, the convex structure is, in a sense, invertible. If we fix $\ens$, $\ens[a]$ and $p$ then, if it exists, there is only one $\ens[b]$ such that $\ens = p \ens[a] + (1-p) \ens[b]$. The ensemble space is \textbf{complement}, meaning that there is only one complement to a component of an ensemble.

However, we found no physical justification for this requirement, and we suspect that it may need to be relaxed in future physical theories.

\subsection{Entropy}
The last requirement is that each ensemble must have a well defined entropy. That is, there is a scalar function $S : \Ens \to \mathbb{R}$ that satisfies the following
\begin{enumerate}
	\item strictly convex: $S(p \ens[a] + (1-p) \ens[b]) - (p S(\ens[a]) + (1-p) S(\ens[b]) ) \geq 0$
	\item upper bounded: $S(p \ens[a] + (1-p) \ens[b]) - (p S(\ens[a]) + (1-p) S(\ens[b]) ) \leq - p \log p - (1-p) \log(1-p)$
\end{enumerate}
The first tells us that, during mixing, the entropy cannot decrease, and it stays the same if and only we are mixing an ensemble with itself. The second tells us that the most the entropy can increase is given by the choice between the two ensembles.

The entropy is ultimately responsible for all geometrical structure in physics. In classical mechanics, the geometry is essentially defined by volumes in phase space. The entropy of uniform distributions is given by the logarithm of the volume, meaning that given the volumes we are able to calculate the entropy and given the entropy we are able to reconstruct the volume. In quantum mechanics, the inner product (i.e. the Born rule) allows us to calculate the entropy of the mixture of two pure states, and given that entropy we can recover the Born rule.

An interesting insight is that the entropy, being strictly concave, has a negative defined Hessian. The negation of the Hessian, then, is a symmetric positive defined function of two infinitesimal variations, and can serve as a metric tensor. In both classical and quantum mechanics this recovers the Fisher-Rao metric.

Another important feature is that if two ensembles saturate the upper bound, then they are orthogonal in both classical and quantum mechanics. Therefore we define two ensembles to be \textbf{orthogonal} if their mixture lead to a maximal entropy increase. If two states are orthogonal, on physical grounds, the must have no common component. However, the reverse is not case: two pure states in quantum mechanics have no components to have in common, but are not necessarily orthogonal. However, classical probability spaces are exactly those space in which no common component implies orthogonality. This is the main source of the ``weirdness'' of quantum mechanics.

To recap, the topological structure captures experimental verifiability and is responsible for the limits and the connection to measureable quantities. The convex structure captures statistical mixing and is responsible for all linear structures of physical theories. The entropic structure captures the variability within an ensemble and is ultimately responsible for all geometric structure. We now concentrate on recovering the measure theoretic structures associated to probability and state counting.

\section{Fraction capacity}

Suppose we have a physical theory that allows for a statistical description. At the very least, this has to provide a notion of ensemble together with the space of all possible ensembles allowed by the theory. We call this an \textbf{ensemble space} $\Ens$. 
Given a target ensemble $\ens \in \Ens$ and an arbitrary ensemble $\ens[a] \in \Ens$ we will be able to write $\ens = p \ens[a] + (1-p) \ens[b]$ for some $p \in [0,1]$ and some $\ens[b] \in \Ens$. At the very least, we can use $p=0$ and $\ens[b] = \ens$. Given that $p$ is bounded, there will be a biggest $p$ possible which we call the \textbf{fraction} of $\ens[a]$ in $\ens$. Mathematically
\begin{equation}
	\fraction_{\ens}(\ens[a]) = \sup(\{ p \in [0,1] \, | \, \exists \, \ens_1 \in \Ens \text{ s.t. }  \ens = p \ens[a] + \bar{p} \ens_1 \}).
\end{equation}
This quantity tells us how much of ensemble $\ens$ can be constructed, can be characterized, by $\ens[a]$.

We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a set of ensembles $A \subset \Ens$., meaning we want to characterize how much of a target ensemble $\ens$ can be constructed as a mixture of ensembles from $A$. Given a set of ensembles $A$, its \textbf{hull}, noted $\hull(A)$ is the set of all mixtures that can be constructed from $A$. That is, is the set of all convex combinations $\sum_i^n p_i \ens[a]_i$ such that $\{a_i\}_1^n \subseteq A$. The \textbf{fraction capacity} is the biggest fraction among all the elements of the hull. Mathematically
\begin{equation}
	\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\}).
\end{equation}

If $\Ens$ is a topological space, we can imagine the fraction capacity to be defined on the sigma algebra $\Sigma_{\Ens}$. We can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
\begin{enumerate}
	\item non-negative and unit bounded - $0 \leq \frcap_{\ens}(A) \leq 1$
	\item monotone - $A \subseteq B \implies \frcap_{\ens}(A) \leq \frcap_{\ens}(B)$
	\item sub-additive - $\frcap_{\ens}(A \cup B) \leq \frcap_{\ens}(A) + \frcap_{\ens}(B)$.
	\item continuous from below and above - $\frcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \frcap_{\ens}(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
\end{enumerate}

\section{Statistical properties}

\section{Beyond real valued quantities}

\section{State capacity}

\section{Quantization}
* 3 pick 2

\section{Quantizing space-time}
* we need a non-additive measure on degrees of freedom
*

\section{Conclusion}

\section*{Acknowledgments}
This paper is part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived. This article was made possible through the support of grant \#62847 from the John Templeton Foundation.

\bibliography{bibliography}

\newcommand{\pj}[1] {\underbar{$#1$}}

\end{document}

### 2024-10-10 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: more work on conceptual part

@@
 
-We are looking to find the minimal set of axioms required by a physical theory that allows statistical descriptions.
-Such a theory must, at the very least, provide us with a notion of ensemble, and a space of ensembles that are allowed in the theory.
+We are looking to find a minimal set of axioms required by a physical theory that allows statistical descriptions.
+Such a theory must, at the very least, provide us with a notion of ensemble, and the space of ensembles that are allowed in the theory.
 An ensemble can be understood as the infinite collection of the outputs of a particular preparation procedure.
@@
 In quantum mechanics, instead, it is the space of density operators.
-While the ensemble space for each theory will be in principle different, some basic features will remain the same.
-We are looking for these basic features, necessary elements that any ensemble space must have.
+While the ensemble space for each theory will be in principle different, some will have many common features.
+In all cases we must be able to talk about conservative or dissipative processes, measurable properties of the system, probability distributions for outcomes of said properties and so on.
+We are looking for necessary minimal basic requirements upon which these common tools can be built.
 
-Note that there may good philosophical grounds to require that a physical theory allows for statistical descriptions.
-If a physical theory has to be testable, and repeatedly so, then we fail to see how it could be so without providing statistical tools.
+Note that there may good conceptual grounds to require that a physical theory allows for statistical descriptions.
+If a physical theory is to be repeatedly testable experimentally, then we fail to see how it could be so without providing statistical tools.
 Every state preparation and measurement, in fact, is typically statistical in nature.
-The development of such arguments would be interesting to us.
+Moreover, physical laws are statements of the type ``whenever we prepare this, we can measure that.'' A philosophical exploration of this area would be welcome.
 
-We have three basic requirements on ensembles: first, they have to be identifiable experimentally; second, they need to allow mixtures; lastly, they need an entropy well defined.
+We have identified three basic requirements on ensembles: first, they have to be identifiable experimentally; second, they need to allow mixtures; lastly, they need a well-defined entropy.
 Let us go through these items one by one.
@@
 \subsection{Experimental verifiability}
+The first requirement is that ensemble must be connected to experimental verification.
+That is, we must have enough \textbf{verifiable statements} at our disposal to define ensembles and tell them apart.
+By verifiable statement we mean a statement for which an experimental test is available that terminates in finite time if and only if the statement is true.
+For example, the statement ``the mass of the photon is less than $10^{-13} \, eV$'' is verifiable, while ``the mass of the photon is exactly $0 \, eV$'' is not verifiable due to its infinite precision.
+Mathematically, this requirement forces an ensemble space to be a $T_0$ second countable topological space, where open sets represent verifiable statements and Borel sets represent testable statements without guarantee of termination.
+
+The topological structure, then, represent the most foundational structure for a physical theory.
+It tells us why functions are ``well-behaved,'' which simply means they have to be topologically continuous in the ``natural topology'' induced by this requirement.
+It tells why the Banach-Tarski paradox does not apply in physics since non-Borel sets are physically ill-defined as they are not connected to experimental verification.
+Similarly, it tells us that sets with cardinality greater than that of the continuum are physically not relevant, as they cannot be given a $T_0$ second countable topology.
 
@@
 
+\subsection{Interplay between structures}
+
+Each structure is important, but the balance between them is even more important.
+The key is to understand, on physical grounds, where each problem lies and therefore use the appropriate tool.
+For example, one may focus too much on the convex/linear structure and may be tempted to close the space under infinite convex/linear combinations.
+Mathematically, it would be very convenient.
+Physically, it is unsound.
+It is the link to experimental verification that tells us which limits are allowed, that particles cannot have an infinite average position.
+Therefore, it is the topology that has to tells us which limits are allowed, not the convex structure.
+Conceptual clarity has to come first.
+The math follows.
+A full examination of these issues is beyond the scope of this article, and we refer to our larger, and future, work.
+However, there is one key issue that we need to discuss as, we found, it was the main conceptual obstacle to progress.
+
+On the convex structure we can define \textbf{pure states} as the ones that cannot be decomposed into other ensembles.
+Mathematically, these are the extreme points of the space and all other points can be understood as convex combinations, probability distributions, over them.
+So, for example, the pure states for a six-sided die are the six possible outcomes, and all ensembles are simply convex combinations $\sum_{i+1}^{6} p_i = 1$.
+This is exactly how classical discrete spaces work.
+As physicists, we are somehow trained to think that that's how everything works.
+In classical mechanics the pure states are perfect preparations of position and momentum.
+In a classical field theory, they are perfect preparation of all fields at each point in space.
+Things start breaking down in quantum mechanics, where we do have pure states, but then we also want pure states of position and momentum, which kind of work if you are careful, though we typically aren't.
+And then there is quantum field theory.
+
+If we look closely, this does not make sense on physical grounds. The fact that the math becomes harder is simply a consequence of trying to fit a square peg in a round hole.
+
+Perfect preparation of classical particles, continuous quantities in quantum mechanics and field configurations in any field theory are not things we can actually achieve.
+Physically, all ensembles we can prepare have finite precision and finite entropy, and those preparations have neither.
+They are not ensembles and they are not part of the a convex structure.
+In the classical case, we can understand them as infinitesimal decompositions: convex combinations that have more and more terms.
+This is true, but we cannot do that in the quantum case.
+If we have a pure state, this cannot be decomposed.
+Therefore an eigenstate of position is not a limit of decomposition, a limit in the convex structure, but a limit of a sequence of pure states, a limit in the topology, and the topology will not even include that limit.
+We may be able to see a pure state, in some sense, as a superposition of eigenstates of position, but that structure is only available in quantum mechanics, and it does not have a clear physical correspondence.
+
+The insight is that probability distributions for a quantity are not defined by mixing the best ensembles we can prepare (i.e.
+a convex combination of extreme points), but by assigning a probability over each experimentally verifiable (i.e.
+open) interval of the possible values of said quantity (its spectra).
+The two coincide over a discrete set of orthogonal pure states, which is the least physically interesting case.
+If our are to be physically well defined and true generalizations, they must work in the more complicated case.
+We not only have to be able to talk about probability in terms of mixtures of ensembles, but we must be able to recover the interval of possible values for physical properties (i.e.
+the spectra), the object we typically assign probability two.
+
 To recap, the topological structure captures experimental verifiability and is responsible for the limits and the connection to measureable quantities.
@@
 
-Suppose we have a physical theory that allows for a statistical description.
-At the very least, this has to provide a notion of ensemble together with the space of all possible ensembles allowed by the theory.
-We call this an \textbf{ensemble space} $\Ens$.
+Note that the mixing coefficient $p$ is connected to probability but it is not itself probability.
+It represents a parameter for the mixing states, mixing preparations, while probability is typically understood as the chance of something happening, the likelihood of a final event.
+Mathematically, the probability measure is a function from an event, which can be understood as a collection of fully distinct outcomes, to a number from zero to one.
+If we are mixing ensemble that represent fully distinct outcomes (i.e.
+they are orthogonal), then the mixing $p$ is a probability.
+If we are mixing ensemble that do not represent fully distinct outcomes (e.g.
+spin up and spin left), the probability of getting one of the outcomes is not the mixing coefficient $p$.
+
+This is exactly what happens in quantum mechanics: probability is defined only on a set of orthogonal states.
+If we focus only on preparation, then, we only talk about mixing coefficients across non-orthogonal elements and define a more general tool that, when restricted on orthogonal sets, will recover probability.
+
+We concentrate on the following question: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subset \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
+How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
+For example, if $\ens$ is a uniform distribution over a fair six-sided die, and $A$ contains four ensemble, each fully on ${1,2}$ and uniform distribution over ${5,6}$ two and four, than $5/6$ of $\ens$ can be constructed from $A$, which is the probability that we d
+
 Given a target ensemble $\ens \in \Ens$ and an arbitrary ensemble $\ens[a] \in \Ens$ we will be able to write $\ens = p \ens[a] + (1-p) \ens[b]$ for some $p \in [0,1]$ and some $\ens[b] \in \Ens$.

### 2024-10-11 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: another pass at explaining the core issue

@@
 The key is to understand, on physical grounds, where each problem lies and therefore use the appropriate tool.
-For example, one may focus too much on the convex/linear structure and may be tempted to close the space under infinite convex/linear combinations.
+For example, one may focus on the convex/linear structure and may be tempted to close the space under infinite convex/linear combinations.
 Mathematically, it would be very convenient.
-Physically, it is unsound.
-It is the link to experimental verification that tells us which limits are allowed, that particles cannot have an infinite average position.
-Therefore, it is the topology that has to tells us which limits are allowed, not the convex structure.
-Conceptual clarity has to come first.
+However, not all infinite mixtures correspond to physically meaningful systems, such as a particle with an infinite average position.
+It is the connection to experimental verification, the topology, that will tell us which limits are allowed and which are not.
+This type of conceptual clarity has to come first.
 The math follows.
@@
 On the convex structure we can define \textbf{pure states} as the ones that cannot be decomposed into other ensembles.
-Mathematically, these are the extreme points of the space and all other points can be understood as convex combinations, probability distributions, over them.
-So, for example, the pure states for a six-sided die are the six possible outcomes, and all ensembles are simply convex combinations $\sum_{i+1}^{6} p_i = 1$.
+For example, the pure states for a six-sided die are the ensemble perfectly prepared respectively on the six possible outcomes.
+All ensembles are convex combinations $\sum_{i+1}^{6} p_i = 1$ of those pure states.
+Mathematically, the pure states are the extreme points of the space and all other points can be understood as convex combinations over them.
 This is exactly how classical discrete spaces work.
-As physicists, we are somehow trained to think that that's how everything works.
-In classical mechanics the pure states are perfect preparations of position and momentum.
-In a classical field theory, they are perfect preparation of all fields at each point in space.
-Things start breaking down in quantum mechanics, where we do have pure states, but then we also want pure states of position and momentum, which kind of work if you are careful, though we typically aren't.
-And then there is quantum field theory.
+Most physicists, including are older selves, think that that's how everything works.
+For example, in classical mechanics the pure states are perfect preparations of position and momentum, and ensembles are probability distributions.
+This does not work.
 
-If we look closely, this does not make sense on physical grounds. The fact that the math becomes harder is simply a consequence of trying to fit a square peg in a round hole.
+We cannot consistently prepare classical particles perfectly, therefore the points of phase space are not ensemble.
+The mathematical facts that delta Dirac are not proper functions, have support with empty interior, have minus infinite entropy, and so on, are simply a consequence of our conceptual problem.
+A bit better is to think them as the limit of infinitesimal decomposition: classical ensembles can be decomposed into more and more precise ones, and the points of are the limit of this decomposition.
+Probability distributions are defined on these limits.
+This works in classical mechanics, but in general.
 
-Perfect preparation of classical particles, continuous quantities in quantum mechanics and field configurations in any field theory are not things we can actually achieve.
-Physically, all ensembles we can prepare have finite precision and finite entropy, and those preparations have neither.
-They are not ensembles and they are not part of the a convex structure.
-In the classical case, we can understand them as infinitesimal decompositions: convex combinations that have more and more terms.
-This is true, but we cannot do that in the quantum case.
-If we have a pure state, this cannot be decomposed.
-Therefore an eigenstate of position is not a limit of decomposition, a limit in the convex structure, but a limit of a sequence of pure states, a limit in the topology, and the topology will not even include that limit.
-We may be able to see a pure state, in some sense, as a superposition of eigenstates of position, but that structure is only available in quantum mechanics, and it does not have a clear physical correspondence.
+In quantum mechanics we have pure states, and these are ensembles that cannot be further decomposed into finer ensembles.
+Therefore an eigenstate of position is not a limit of decomposition, a limit in the convex structure.
+It is the limit where we are preparing a pure state on a smaller and smaller region of space.
+It is the limit of a sequence of pure states, a limit in the topology.
+Physically, more than mathematically, it is a completely different thing, therefore it is not a good route to generalization.\footnote{The fact that pure states can be seen, in some sense, as superposition of eigenstates of position, makes the matter even more confusing.}
 
-The insight is that probability distributions for a quantity are not defined by mixing the best ensembles we can prepare (i.e.
+The insight is that probability distributions for a quantity are not defined by statistical mixtures of perfect preparations (i.e.
 a convex combination of extreme points), but by assigning a probability over each experimentally verifiable (i.e.
 open) interval of the possible values of said quantity (its spectra).
-The two coincide over a discrete set of orthogonal pure states, which is the least physically interesting case.
-If our are to be physically well defined and true generalizations, they must work in the more complicated case.
-We not only have to be able to talk about probability in terms of mixtures of ensembles, but we must be able to recover the interval of possible values for physical properties (i.e.
-the spectra), the object we typically assign probability two.
+The two coincide in classical discrete space, which is the least physically interesting case as it does not even allow for classical particle mechanics.
 

### 2024-10-30 Articles/2023-HilbertQuantum/Submissions/Quantum Studies/HilbertQuantum.tex
HilbertQuantum: adding extra section for motivation

@@
 
+\section{On physicality}
+
+When we first drafted this paper, we naively thought that the advantage of having a well-posed mathematical structure that represents, and only represents, physically well-defined objects would be self-evident.
+It does not seem to be the case: a previous reviewer stated that ``the question of whether the limiting states are actually physical is rather pointless;'' conversely, others see no problem in working with QFT path integrals that are not mathematically well-defined.
+We are therefore in the absurd, from our perspective, position of having to defend that, yes, the math we need to use in physics has be mathematically precise and, yes, the math we need to use in physics has to represent, perhaps idealized, physically realizable objects.
+The cost of this failure is all around us: an ever increasing number of interpretations of quantum mechanics, abstract mathematical work whose connection to physics is unclear, theoretical physics more and more disconnected from experimental physics.
+The reader that, like us, finds this obvious can skip this section.
+
+One problem is that most physicists see mathematics as just a tool for calculation, much like spectroscopes or telescopes are tools to gather data.
+As long as they do what they are supposed to do, the details are irrelevant because the physics lies somewhere else.
+If they do not work correctly, it is just a detail for the expert to fix.
+But mathematics plays a much more fundamental role in physics: it is the language in which we specify our models.
+When we say ``states are rays in a Hilbert space'' we are giving a physical specification.
+We are saying that there are objects in the real world that can be modeled with a particular mathematical structure.
+If Hilbert spaces do not work, it is not simply a math problem: the model is wrong, which is a physics problem.
+This is the question we are examining in this paper: does the math we use in quantum mechanics model the physical world correctly?
+Is our physical problem well specified?
+
+To make this determination, we need to provide a standard to judge the physicality of a formal definition.\cite{redei_tension_2020, north_physics_2021} Suppose we are studying a physical system.
+A mathematical definition is \textbf{physical} if it properly characterizes the physical system in the given conditions.
+That is, if we are able to \textbf{justify on physical grounds} that the given mathematical definition is needed to capture and only capture a particular aspect of the physical system under the given condition.
+On the other hand, a mathematical definition is \textbf{unphysical} if it can be shown to require properties or operations that cannot have a physical counterpart.
+The question of this paper, then, is whether the use of Hilbert spaces to characterize quantum systems can be justified on physical grounds, from physical requirements.
+This is the task at hand.
+
+Some mathematicians were disappointed that we didn't discuss in depth approaches like rigged Hilbert spaces or the algebraic approach to quantum mechanics.
+Do rigged Hilbert spaces spell out exactly what each element of the Gelfand triple describes and give a physical justification that tells us exactly which of those are the physically correct specifications?
+Do algebraic approaches give a physical justification as to why things we measure in a lab must be modeled by a $C^*$-algebra, and consequently allow us to determine whether the use of Hilbert spaces to model states can be given (or not given) a physical justification?
+To our knowledge, and to the knowledge of the mathematicians that approached us, they do not.
+This does not mean that these works are of no interest to a physical audience.
+It just means they are not going to help with the task at hand.
+
+While it seems obvious to us, it may be worth stating that \textbf{it is not up to mathematicians to justify the use of a particular mathematical structure in physics}.
+Ultimately, the justification cannot be solely in terms of mathematical matters, but rather it has to rely on that elusive ``physical intuition.'' In fact, the task is exactly to make that elusive ``physical intuition'' precise enough to turn it into mathematical definitions.
+This is the job of a physicist.
+In our specific case, the core argument we will use is that states described by measurable quantities with infinite or ill-defined expectation values are not physically realizable.
+To us, and to every experimental physicist we talked to, ``it doesn't make sense experimentally,'' and this is the best argument possible to show that something is unphysical.
+``It can't be done.'' No argument is better.
+However, more than one previous reviewer found no problem with infinite (or undefined) expectation values, as they adopt an ``ensemble interpretation,'' and tried to convince us that we shouldn't either.
+How can progress on the mathematical foundations of quantum mechanics be made if an argument that is accepted by the vast majority of physicists is rejected?
+
+Obviously, while the argument is ultimately physical, and it is therefore the job of a physicist to make these arguments, it has to be a physicist that is well-versed in foundational issues in mathematics.
+The justification, in fact, will need precise understanding of the mathematical definitions and their implication for both the physics and the math.
+Let us give an example.
+In a previous work we found that a set of physically distinguishable objects must be a $T_0$ second countable topological space where each open set represents a verifiable statement, a statement allowing a test that terminates in finite time if and only if the statement is true.
+This result tells us why functions in physics must be ``well-behaved'' (i.e.~topologically continuous in the right topology), why sets with cardinality greater than the continuum are unphysical (i.e.~they cannot be given a $T_0$ second countable topology) and why the Banach-Tarski paradox does not apply in physics (i.e.
+non-Borel sets are unphysical as they cannot be associated with an experimental test).
+This tells us that, at the very least, non-separable Hilbert spaces are unphysical as they have cardinality greater than the continuum.\cite{pittphilsci18363} This ties together ideas from point-set topology, logic, set theory, measure theory and computer science.
+Clearly, one needs to have at least a basic understanding of these fields to be able to follow the argument, and most degree programs in physics do not include typically discuss these topics in depth.\footnote{A previous reviewer conflated the completeness of the Hilbert space with the existence of a complete orthonormal basis of said space.} However, in the end, only a physicist can judge whether the original premise, the evidence-based requirement of physics, justifies the mathematical requirements.
+
+To summarize, we are not interested in understanding whether Hilbert spaces have nice mathematical features for calculating or proving results.
+The fact that other tools have been developed to solve other mathematical problems is not relevant to this discussion.
+We are only interested in the question of mathematical representation of quantum states, \textbf{physical} entities we prepare in a lab.
+Something we produce, manipulate and, in the end, measure.
+What physical requirements are we capturing when we model them mathematically?
+Are Hilbert spaces capturing, and only capturing, physically tenable assumptions, or not?
+
 \section{On the physicality of Hilbert spaces}
 
-Let us first clarify the standard we believe should be used to judge the physicality of a formal definition.\cite{redei_tension_2020, north_physics_2021} Suppose we have a physical system under study in a given condition.
-A mathematical definition is \textbf{physical} if it properly characterizes the physical system.
-That is, if it can be shown that the mathematical definition is needed to capture and only capture a particular aspect of the physical system under the given condition.
-On the other hand, a mathematical definition is \textbf{unphysical} if it can be shown to require properties or operations that cannot have a physical counterpart.
-
-For example, suppose we have a set of physically distinguishable elements (e.g.
-states, bodies, regions of space), meaning that we can tell any two apart through an experimental test.
-Experimentally, the best we can do is collect finite information in finite time.
-If given an arbitrarily long time, we can collect an arbitrarily large but finite amount of information.
-An element of the set, then, must be identified by, at most, a countably infinite sequence of ones and zeros so that, at some point, we can tell the difference between the two sequences and, therefore, the elements.
-The set of countable sequences of zeros and ones can be understood as the binary expansion of real numbers between zero and one, which has the cardinality of the continuum.
-Therefore, any mathematical definition requiring a set of physical elements to have cardinality greater than the continuum is \textbf{unphysical}.
-For example, non-separable Hilbert spaces have cardinality greater than the continuum and, therefore, are unphysical.\cite{pittphilsci18363}
-
-Naturally, this does not mean all sets with a cardinality of the continuum or lower are physical.
-In general, testing physicality is harder as we need to check that all the mathematical features map to physical features and that no physical feature is left behind.
-Clearly, postulating that ``the state space of quantum states has cardinality no greater than that of the continuum'' is certainly \textbf{physical}, but this tells us little.
-Let us then see what features of Hilbert spaces we should consider physical in the context of quantum mechanics.
+As we have discussed in the previous section, to argue for the physicality of a mathematical definition we need to make a justification \textbf{on physical grounds} that the definition captures a physical property.
+We will see that the inner product space part of the Hilbert space definition is physical as it is required to to compute entropy and probabilities.
+We have also seen that to argue for the unphysicality of a mathematical definition we need to show that the definition implies objects that are physically unrealizable.
+We will see that the completeness part of the Hilbert space definition is unphysical as it forces us to include states with infinite (or undefined) expectation values.
 

### 2024-11-05 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Intro and first section review

@@
 \section{Introduction}
-Paper for physicists:
-* argue that we need to develop mathematical structures from physical requirements
-* create a theory of probability that works for classical, quantum mechanics and beyond
-* inverted structure
+
+Our current understanding of probability is given in terms of the measure theoretic formalization due to Kolmogorov, and its various philosophical interpretations.
+However, probability in quantum mechanics, as it is widely known, is not Kolmogorovian, in the sense that it provides results that are not reproducible in standard probability spaces.
+It follows that our mathematical axioms for probability and their interpretations are insufficient for physical theories.
+
+An alternative measure theoretic strategy, that has been successful is quantum mechanics, is the use of quasi-probability distributions: measures that are still unitary over the whole space, but can have negative values in small regions.
+However, these haven't resulted in an alternative axiomatic approach with suitable interpretations for at least, in our mind, three reasons.
+First, it is not clear how negative probability should be interpreted.
+Second, it is not clear how to characterize the space of valid quasi-probability distributions (e.g.
+there is no simple rule to specify which $\rho(q,p)$ are valid Wigner functions).
+Third, there are multiple options for quasi-probability distributions (e.g.
+Wigner function, Husimi distribution, Sudarshan–Glauber distribution).
+This lack of foundational clarity presents an additional problem: how do we know whether future physical theories will not require yet another extension?
+
+The goal of this paper is to show that, in the same way that classical probability can be axiomatized through standard measure theory, quantum probability, and in fact any probability used a physical theory, can be axiomatized through non-additive (i.e.
+fuzzy) measure theory.
+We will define a generalized concept of probability that is ultimately based on physically motivated axioms, giving the physicist a clear operational motivation to our definitions, the mathematicians a formally well-defined mathematical structure and the philosopher a unified framework in which probability can be interpreted in the context of physical theories.
+We are not going to be able to develop a full theory here, as there are still a number of conceptual and technical challenges to be solved that will require work in physics, mathematics and philosophy.
+However, we will show that there are enough results that point in a good direction, and therefore make it reasonable to invest time and effort to work on the open problems.
+For those interested, the current state of the research will always be available in an open document that will be routinely updated as the theory matures.
+
+In a nutshell, we will start by arguing that physical theory must at least be able to talk about statistical ensembles, and that ensembles must allow the preparation of statistical mixture.
+Given a target ensemble $\ens$ and a set of ensemble $A$, we define the fraction capacity of $A$ for $\ens$ as the biggest fraction of $\ens$ that can be constructed from another set of ensembles $A$.
+That is, the highest $p$ such that we can express our ensemble as $\ens = p \ens[a] + (1-p) \ens[b]$, where $\ens[a]$ is a mixture of elements of $A$.
+For example, if $\ens$ is the uniform distributions for a roll of a six-faced die, and $A$ includes only two ensembles with $100\%$ probability for the outcomes $3$ and $4$ respectively, then the faction capacity of $A$ for $\ens$ will be $\frac{1}{3}$.
+Note that this coincides with the probability of the event ``$3$ or $4$''.
+The fraction capacity is a non-negative, unit bounded, sub-additive and continuous set function, which is therefore like a probability measure except for the additivity, which is recovered for classical mechanics and over quantum measurements.
+
+\section{Quick review of probability theory}
 
 \section{Ensemble spaces}
+
+As mentioned in the introduction, our goal is to construct a probability theory that is guaranteed to work in any physical theory.
+Therefore our fundamental axioms will not be about probability, but about physical theories.
+This will help us place probability in a bigger context, and provide a minimal interpretation.
+As it is common in our larger project, we want axioms to describe necessary requirements of the scientific practice.
+That is, they will describe features of physical practice, not feature of the systems being described.
 
@@
 
-Note that there may good conceptual grounds to require that a physical theory allows for statistical descriptions.
-If a physical theory is to be repeatedly testable experimentally, then we fail to see how it could be so without providing statistical tools.
-Every state preparation and measurement, in fact, is typically statistical in nature.
-Moreover, physical laws are statements of the type ``whenever we prepare this, we can measure that.'' A philosophical exploration of this area would be welcome.
+At least two arguments can be made to show that any physical theory must provide a set of ensembles.
+First, note that physical laws are never about specific instances but rather regularities.
+They are statements of the type ``whenever we prepare this, we can measure that.'' This means that, ultimately, they are about ensembles.
+Second, if a physical theory is to be testable experimentally, and repeatedly so, it must be in terms of ensembles because this is what we prepare in practice and this is what, in the end, we characterized experimentally.
+It seems, then, that the approach can be given firm conceptual grounds, which we leave as open but solvable philosophical problem.
 
-We have identified three basic requirements on ensembles: first, they have to be identifiable experimentally; second, they need to allow mixtures; lastly, they need a well-defined entropy.
+We have identified three basic requirements on ensembles: first, they have to be identifiable experimentally; second, they need to allow statistical mixtures; lastly, they need a well-defined entropy.
 Let us go through these items one by one.
+We leave the full details to the working document (TODO: decide how to call it).
 
@@
 For example, the statement ``the mass of the photon is less than $10^{-13} \, eV$'' is verifiable, while ``the mass of the photon is exactly $0 \, eV$'' is not verifiable due to its infinite precision.
-Mathematically, this requirement forces an ensemble space to be a $T_0$ second countable topological space, where open sets represent verifiable statements and Borel sets represent testable statements without guarantee of termination.
+Based on previous work, this mathematically requires an ensemble space to be a $T_0$ second countable topological space, where open sets represent verifiable statements and Borel sets represent statements associated with a test, but without guarantee of termination.
 
@@
 It tells us why functions are ``well-behaved,'' which simply means they have to be topologically continuous in the ``natural topology'' induced by this requirement.
-It tells why the Banach-Tarski paradox does not apply in physics since non-Borel sets are physically ill-defined as they are not connected to experimental verification.
+It tells why probability is assigned to Borel sets (i.e.
+all statements associated with tests) and why the Banach-Tarski paradox does not apply in physics (i.e.
+non-Borel sets are physically ill-defined as they are not connected to experimental verification).
 Similarly, it tells us that sets with cardinality greater than that of the continuum are physically not relevant, as they cannot be given a $T_0$ second countable topology.
@@
 \subsection{Statistical mixtures}
-Another basic requirement of an ensemble space is ability to prepare a \textbf{mixture} of two ensembles.
-If $\ens[a], \ens[b] \in \Ens$ are two ensembles, and $p \in [0,1]$ a weight, we will find the $\ens = p \ens[a] + (1-p) \ens[b] \in \Ens$ that describes a process that selects the first ensemble over the second $p$ percent of the times.
+Another basic requirement is the ability to prepare a \textbf{mixture} of two ensembles.
+If $\ens[a], \ens[b] \in \Ens$ are two ensembles, and $p \in [0,1]$ a weight, then $\ens = p \ens[a] + (1-p) \ens[b] \in \Ens$ that describes a process that selects the first ensemble over the second $p$ percent of the times.
 Mathematically, the ensemble space is endowed with a \textbf{convex structure}.
@@
 If we fix $\ens$, $\ens[a]$ and $p$ then, if it exists, there is only one $\ens[b]$ such that $\ens = p \ens[a] + (1-p) \ens[b]$.
-The ensemble space is \textbf{complement}, meaning that there is only one complement to a component of an ensemble.
+The ensemble space is \textbf{complemented}, meaning that there is only one complement to a component of an ensemble.
 
-However, we found no physical justification for this requirement, and we suspect that it may need to be relaxed in future physical theories.
+This perspective allows to understand what negative probability is.
+Suppose we have a real vector space of dimensions $n$, then $n$ linearly independent vectors are enough to express every point as a linear combination.
+If we add another point, we can now express every point as an affine combination, that is a linear combination where the coefficients sum to one, though some can still be negative.
+Since the space of mixed space in quantum mechanics is a real vector space, we can choose a linearly dependent set of states space span the whole space, and every mixed state can be expressed as an affine combination, as a pseudo-probability.
+
+Pseudo-probability distributions, then, require the additional vector space structure.
+However, we found no physical justification to require the additional structure, and we suspect that it may need to be relaxed in future physical theories.
+Therefore it is important that our fundamental definitions require only the convex structure, and not the vector space structure.
 
@@
 The first tells us that, during mixing, the entropy cannot decrease, and it stays the same if and only we are mixing an ensemble with itself.
-The second tells us that the most the entropy can increase is given by the choice between the two ensembles.
+The second tells us that the most the entropy can increase is given by the choice between the two ensembles.\footnote{The use of the Shannon entropy for the upper bound can likely be recovered.}
 

### 2024-11-05 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
ClassicalLimit: first arXiv version

@@
 \documentclass{article}
+\usepackage{authblk}
 
@@
 \usepackage[english]{babel}
-
+\usepackage{ulem}
 % Set page size and margins
@@
 \usepackage{assumptionsofphysics}
+\usepackage[dvipsnames]{xcolor}
 \usepackage{tikz-cd}
@@
 \usepackage{pgfplots}
+
 \def\>{\rangle}
@@
 
-\title{Classical mechanics as high entropy limit}
-\author{Gabriele Carcassi, Manuele Landini, Christine A. Aidala}
+\title{Classical mechanics as the high-entropy limit \\ of quantum mechanics}
+\author[1]{Gabriele Carcassi}
+\author[2]{Manuele Landini}
+\author[1]{Christine A. Aidala}
+
+\affil[1]{Physics Department, University of Michigan, Ann Arbor, Michigan 48109, USA}
+
+\affil[2]{Institut f\"ur Experimentalphysik und Zentrum f\"ur Quantenphysik, \protect\\ Universit\"at Innsbruck, Innsbruck, Austria}
 
@@
 \begin{abstract}
-We show that classical mechanics can be recovered as the high entropy limit of quantum mechanics.
-The idea is that the high entropy hides quantum effects, and mixed states of high enough entropy can be approximated with classical distributions.
+We show that classical mechanics can be recovered as the high-entropy limit of quantum mechanics.
+That is, the high entropy masks quantum effects, and mixed states of high enough entropy can be approximated with classical distributions.
 The mathematical limit $\hbar \to 0$ can be reinterpreted as setting the zero entropy of pure states to $-\infty$, in the same way that non-relativistic mechanics can be recovered mathematically with $c \to \infty$.
 Physically, these limits are more appropriately defined as $S \gg 0$ and $v \ll c$.
-Both limits can then be understood as approximations independently of what mechanism allows those approximations to be valid.
-Consequently, the limit presented is independent of what interpretation is chosen for both quantum states and entropy.
+Both limits can then be understood as approximations independently of what circumstances allow those approximations to be valid.
+Consequently, the limit presented is independent of possible underlying mechanisms and of what interpretation is chosen for both quantum states and entropy.
 \end{abstract}
@@
 
-%TODO: Status. Probably two ways: one mathematical (hbar to zero) and multiple mechanism (what happens that makes things loook classical) like decoherence and other things
+The goal of this paper, summarized in figure \ref{fig_quadrants}, is to show that classical mechanics is the high-entropy limit of quantum mechanics, much in the same way that it is the low-speed limit of relativistic mechanics.
 
-%TODO: Our aim is to recover the mathematical limit for general physical ideas, in a way that is independent of mechanism and interpretation.
+\begin{figure}[h]
+	\centering
+\begin{tikzpicture}
+    % Box dimensions
+    \def\boxwidth{4}
+    \def\boxheight{2}
 
-%TODO: give brief overview
+    % Draw four boxes
+    \draw (0,0) rectangle (\boxwidth,\boxheight);
+    \draw (-\boxwidth,0) rectangle (0,\boxheight);
+    \draw (0,-\boxheight) rectangle (\boxwidth,0);
+    \draw (-\boxwidth,-\boxheight) rectangle (0,0);
+
+    % Label Boxes
+    \node[align=center] at (\boxwidth/2,\boxheight/2) {Relativistic \\ Mechanics};
+    \node[align=center] at (-\boxwidth/2,\boxheight/2) {Classical \\ Mechanics};
+    \node[align=center] at (-\boxwidth/2,-\boxheight/2) {Quantum \\ Mechanics};
+    \node[align=center] at (\boxwidth/2,-\boxheight/2) {Quantum \\ Field Theory};
+
+    % Axes
+    \draw[->] (-\boxwidth,\boxheight+0.5) -- (\boxwidth,\boxheight+0.5) node[above,midway] {Speed};
+    \draw[->] (-\boxwidth-0.5,-\boxheight) -- (-\boxwidth-0.5,\boxheight) node[above,midway,rotate=90] {Entropy};
+
+\end{tikzpicture}
+	\caption {The amended four-quadrant picture often used to compare the theories.
+The distinction between classical and quantum physics is not size, but entropy.
+Small systems in states of high entropy (e.g.~a few molecules at high temperature) can be described by classical mechanics; large systems in states of low entropy (e.g.~systems entangled over long distances) cannot be described by classical mechanics.
+} \label{fig_quadrants}
+\end{figure}
+
+Physically, as we take states at higher and higher entropy, the quantum features become less and less pronounced to the point that classical mechanics becomes a good approximation.
+This limit is therefore independent of what interpretation one may give to either quantum mechanics or entropy, and from the reason, the mechanism, that makes a high-entropy description suitable.
+Quite literally, quantum mechanics is needed exactly when we need more precise (i.e.~low-entropy) descriptions.
+This brings the classical limit in line with the non-relativistic limit, where the low-speed limit is seen as unrelated to discussions about the nature of space-time and no one tries to find a reason, a mechanism, as to why speeds become low in certain regimes.
+
+Mathematically, it is common to see both limits as group contractions.
+This too can be given a physically meaningful interpretation.
+The low-speed limit is mathematically recovered by taking the highest possible speed and setting it to infinity.
+This way, all finite velocities are small compared to the highest possible speed.
+This corresponds to taking $c \to \infty$.
+Conversely, the high-entropy limit is mathematically recovered by taking the lowest possible entropy, zero in quantum mechanics, to minus infinity.
+Note that entropy is the logarithm for the count of states, which is quantified in units of $\hbar$.
+Therefore, taking $\hbar \to 0$ is equivalent to setting the entropy for pure states to minus infinity.
+This way, all finite entropies are large compared to the lowest possible entropy.
+In a nutshell, this is what we will show in section four.
+Since physically it does not make sense to take limits of physical constants, $c \to \infty$ should be understood as $v \ll c$, $\hbar \to 0$ should be understood as $S \gg 0$.
+
+The key insight is that \textbf{high entropy is the one and only requirement}.
+We do not require a particular interpretation of quantum mechanics, of entropy or a particular mechanism responsible for the high entropy.
+Any will do.\footnote{The only requirement for entropy is that it matches the experimentally measured one and that its maximization corresponds to the correct experimental quantities.
+In this vein, the Shannon/Gibbs entropy of classical mechanics and the von Neumann entropy of quantum mechanics are simply the tools we use to make predictions within the respective theories.
+The Boltzmann entropy coincides with the Shannon/Gibbs entropy for a uniform distribution.
+}
+
+The paper is structured as follows.
+In the second section we give intuitive insights as to why the limit makes sense on physical grounds.
+We will argue that to produce quantum effects one does effectively prepare, in one way or another, states at low entropy.
+We will study the relationship between uncertainty and entropy in both classical and quantum mechanics, and see how closely these coincide even with an uncertainty of just a few $\hbar$.
+Lastly, we will see how quantum states at high entropy become closer to each other, and how uncertainty on quantum features can become equivalent to uncertainty on classical features.
+This aliasing is what makes the limit possible.
+In the third section we will give an example of how traditional limits can be reinterpreted as high-entropy limits.
+In the fourth section we will show how the high-entropy limit, mathematically, is like taking $\hbar \to 0$.
+We will study entropy-increasing transformations in classical mechanics and see that these are stretching operations over phase space.
+We then find quantum analogues and note that their effect is to rescale the commutator between position and momentum by a constant.
+Mathematically, this can be understood as rescaling $\hbar$.
+Finally, to give a more physically meaningful insight to the limit, we look at how the Wigner function changes as the phase space is stretched and the entropy increases.
 
@@
 
-As mentioned in the introduction, we believe that classical mechanics should be recovered from quantum mechanics in a way that makes intuitive sense to all those that routinely work with quantum systems.
-Therefore, let us start with a few qualitative considerations that hint at the connection between classical mechanics and high entropy.
+Since we aim to provide an approach to the classical limit that makes intuitive sense to all those that routinely work with quantum systems, we will start with a few qualitative considerations that hint at the connection between classical mechanics and high entropy.
 
 \subsection{Producing quantum states}
-
-TODO: qui sarebbe bello mettere un sacco di esempi/citazioni
 
@@
 Coherence is probably one of the direct and most important properties.
-It has been shown that coherence can be maintained over long distances and among large number of constituents, meaning that quantum systems are not necessarily small or made of few components.
-However, it is also established how coherence can be quickly lost through interaction with the environment, trough decoherence.
-Since decoherence does increase the entropy on the system, it represents one mechanism to reach the high entropy limit.
+It has been shown that coherence can be maintained over long distances and among a large number of constituents, meaning that quantum systems are not necessarily small or made of few components.
+However, it is also established how coherence can be quickly lost through interaction with the environment, through decoherence.\cite{RevModPhys.75.715, SCHLOSSHAUER20191} Since decoherence increases the entropy of the system, it represents one mechanism to reach the high-entropy limit.
 This is in line with our result.
 
-In experimental practice, many quantum effects (e.g.~superconductivity, topological insulators, quantum hall effect, ...) are harder or impossible to achieve at high temperature.
-The thermal noise can in fact break the coherence of the system.
-Note that entropy is a monotonic function of temperature, meaning that decreasing temperature does mean decreasing entropy of the system.
+In experimental practice, many quantum effects (e.g.~superconductivity, topological insulators, quantum Hall effect, ...) are harder or impossible to achieve at high temperature.
+The thermal noise can break the coherence of the system.
+Note that entropy is a monotonic function of temperature, meaning that decreasing temperature means decreasing entropy of the system.
 This is in line with our result.
@@
 Some quantum effects can be replicated at higher temperature given a high pressure.
-This is the case, for example, in some superconductive materials.
-High pressure corresponds to low entropy, which is in line with our result.
+This is the case, for example, in some superconductive materials.\cite{Drozdov:2015} High pressure corresponds to low entropy, which is in line with our result.
+Along the same lines, white dwarfs and neutron stars exhibit quantum effects at high pressure, despite high temperatures.
 
-To produce Bose-Einstein condensates, one needs a high density in phase space, which means both low entropy (to decrease the position spread) and low temperature (to decrease the momentum spread).
+To produce Bose-Einstein condensates, one needs a high density in phase space, which means both high spatial density (to decrease the position spread) and low temperature (to decrease the momentum spread).
+The phase space density (PSD) is directly related to entropy and dimensionality, as exemplified by the famous Sackur-Tetrode equation for the entropy of the ideal gas.
 This, again, is in line with our result.
-{\color{blue} Something of interest: For a trapped condensate, the PSD at the center and S are related by $PSD=e^{5/2+\gamma-S/N}$, where $\gamma$ is the virial coefficient for the trapping potential (Pinske et al Phys.Rev.Lett.
-78, 990 (1997)).
-This has been used experimentally to condense at constant entropy, by changing the shape of the potential (D.
-M.
-Stamper-Kurn, Phys.
-Rev.
-Lett.
-81, 2194 (1998))}
+
+For a trapped condensate, the PSD at the center and entropy $S$ are related by $\text{PSD}=e^{5/2+\gamma-S/N}$, where $N$ is the number of particles and $\gamma$ is the virial coefficient for the trapping potential.\cite{PhysRevLett.78.990} This has been used experimentally to condense at constant entropy, by changing the shape of the potential.\cite{PhysRevLett.81.2194} Note that, however, the system is inhomogeneous, so the condensation happens because entropy is transferred from one part of the system to the other.
+
+For thermodynamic systems in equilibrium, entropy can be calculated directly based on occupation of the quantum states of the system.
+For high temperatures, the occupations are all small and classical statistical mechanics applies.
+Quantum effects related to indistinguishability are found whenever occupations approach unity.
+The entropy per particle drops in the same limit, again in line with our result.
+Note, however, that while infinite entropy corresponds to infinite temperature, the reverse is not necessarily true.
+In systems with finitely many possible states we can find the stationary point of $S$ as a function of $E$ for which $1/T=\partial S/\partial E$ is zero.
+This stationary point represents the maximum entropy reachable by the system, and therefore there is no high entropy limit.
+This is exactly the case where temperature can have negative values.
+Given that the entropy is a concave function of the energy, systems that admit a classical description are exactly those for which no such stationary point exists and temperature cannot be negative or infinite.%\footnote{{\color{blue}We can define temperature, based on the equation $1/T=\partial S/\partial E$.
+In systems with a finite number of degrees of freedom, we can find the stationary point of $S$ as a function of $E$.
+This is because both the minimum and maximum energy states will have $S=0$ or minimal.
+The maximum entropy state should correspond to infinite temperature.
+As a consequence, any microstate becomes equiprobable and the entropy takes the max value $S=k_B\log n_{dof}$} I reworked this point to again show how finitely many possibility does not give us a classical limit.
+}
 
@@
 
-Another hint of the link between high entropy and classical mechanics comes from the relationship between entropy and uncertainty.
+Another qualitative argument that shows the link between high entropy and classical mechanics comes from the relationship between entropy and uncertainty for a single degree of freedom.
 
-In both classical and quantum mechanics, Gaussian states maximize entropy at a fixed uncertainty or, equivalently, minimize uncertainty at fixed entropy.
-This means that, if we fix the entropy $S$, we will have an uncertainty relationship
+In both classical and quantum mechanics, Gaussian states maximize entropy for a single independent degree of freedom at a fixed uncertainty or, equivalently, minimize uncertainty at fixed entropy.
+This means that, if we fix the entropy $S$, we will have an uncertainty relationship\cite{Carcassi:2022bpm}
 \begin{equation}
@@
 \end{equation}
-The specific value of $\Sigma$ will depend on the entropy and on whether we are using classical or quantum mechanics, though the relationship will always be saturated by Gaussian states with no correlation between $x$ and $p$.
+The specific value of the uncertainty $\Sigma$ will depend on the entropy and on whether we are using classical or quantum mechanics, though the relationship will always be saturated by Gaussian states with no correlation between $x$ and $p$.
 
@@
 ]
-\addplot[blue,samples=150,domain=0.1:8.5] {ln(x)+1};
+\addplot[thick,blue,samples=150,domain=0.1:8.5] {ln(x)+1};
 \addlegendentry{\(\text{classical}\)}
-\addplot[red,samples=150,domain=0.5:8.5] {(x+1/2)*ln(x+1/2)-(x-1/2)*ln(x-1/2)};
+\addplot[thick,BrickRed,densely dotted,samples=150,domain=0.5:8.5] {(x+1/2)*ln(x+1/2)-(x-1/2)*ln(x-1/2)};
 \addlegendentry{\(\text{quantum}\)}
-\addplot[green,samples=150,domain=0.5:8.5] {(ln(x)+1) -((x+1/2)*ln(x+1/2)-(x-1/2)*ln(x-1/2))};
+\addplot[thick,ForestGreen,dashed,samples=150,domain=0.5:8.5] {(ln(x)+1) -((x+1/2)*ln(x+1/2)-(x-1/2)*ln(x-1/2))};
 \addlegendentry{\(\text{difference}\)}
@@
 \end{tikzpicture}
-    \caption{Entropy $S$ in nats as a function of uncertainty $\Sigma$ in units of $\hbar$.
-Uncertainty is measured in units of $\hbar$.
-In blue, the classical case $\ln(\Sigma) + 1$.
-In red, the quantum case $\left( \Sigma + \frac{1}{2} \right) \ln \left( \Sigma + \frac{1}{2} \right) - \left( \Sigma - \frac{1}{2} \right) \ln \left( \Sigma - \frac{1}{2} \right)$.
-In green the difference betwen the two.}
+    \caption{Entropy $S$ in nats for a Gaussian state as a function of uncertainty $\Sigma$, measured in units of $\hbar$.
+In solid blue, the classical case $S = \ln(\Sigma) + 1$.
+In dotted red, the quantum case $S = \left( \Sigma + \frac{1}{2} \right) \ln \left( \Sigma + \frac{1}{2} \right) - \left( \Sigma - \frac{1}{2} \right) \ln \left( \Sigma - \frac{1}{2} \right)$.
+In dashed green the difference between the two.}
     \label{fig:uncertainty}
@@
 
-In figure \ref{fig:uncertainty} we can see that the classical and quantum cases are very close when the uncertainty is even just a few units of $\hbar$.
+In figure \ref{fig:uncertainty} we can see that the classical and quantum cases are very close even when the uncertainty is just a few units of $\hbar$.
 Things diverge at about two units of $\hbar$: in quantum mechanics the entropy decreases faster and reaches zero at $\hbar/2$, the bound for the Heisenberg uncertainty principle; classical mechanics reaches zero entropy for $\hbar/e$, and then continues in the negative region.
 
-Note that negative entropy is problematic.
-It directly contradicts the third law of thermodynamics, but we can see other, less axiomatic, problems.
-Given that entropy is additive under composition of independent systems, in the negative region the entropy decreases as the number of particles increases.
-This means that, at low temperature, zero particles give the highest entropy.
-This is one problem with classical theory that is corrected by quantum mechanics.
-
-Studying the relationship between uncertainty and entropy, then, reinforces the idea that quantum mechanics is required at low entropy much like relativity is required at high speeds.
+The convergent prediction at high entropy together with the divergent prediction at low entropy, together with negative values for the classical case, reinforces the idea that quantum mechanics is required at low entropy much like relativity is required at high speeds.
 
@@
 The intrinsic geometry of quantum mixed states makes uncertainty over classical or quantum properties look the same.
-It is this aliasing of different preparations that drives the classical limit, as we can see even in a standard two slit experiment setup.
+It is this aliasing of different preparations that drives the classical limit, as we can see even in a standard two-slit experiment setup.
 
@@
 These identify a two-state system, a qubit.
-We can now imagine equal superpositions of the two states, $\frac{1}{2}|L\> + \frac{1}{2} e^{-\imath \theta} |R\>$ where $\theta$ is the difference in phase between the two components.
-If $\theta$ is zero, we have the state $|+\> = \frac{1}{2}|L\> + \frac{1}{2} |R\>$ and the resulting interference pattern will have a peak in the middle of the screen.
-If $\theta$ is $\pi$, we have the state $|-\> = \frac{1}{2}|L\> + e^{-\imath \pi}\frac{1}{2} |R\>$ and the resulting interference pattern will have a valley in the middle of the screen.
+We can now imagine equal superpositions of the two states, $\frac{1}{\sqrt{2}}|L\> + \frac{1}{\sqrt{2}} e^{-\imath \theta} |R\>$ where $\theta$ is the difference in phase between the two components.
+If $\theta$ is zero, we have the state $|+\> = \frac{1}{\sqrt{2}}|L\> + \frac{1}{\sqrt{2}} |R\>$ and the resulting interference pattern will have a peak in the middle of the screen.
+If $\theta$ is $\pi$, we have the state $|-\> = \frac{1}{\sqrt{2}}|L\> + e^{-\imath \pi}\frac{1}{\sqrt{2}} |R\>$ and the resulting interference pattern will have a valley in the middle of the screen.
 
@@
 
-Mathematically, the trace distance $T(\rho,\sigma)$ between two mixed states $\rho$ and $\sigma$ will decrease under a completely positive trace-preserving (CPTP) map $M$ that increases the entropy of all states.
-That is, $T(M(\rho),M(\sigma)) < T(\rho,\sigma)$.\footnote{TODO Man: citation needed} If we call $\mathcal{G}$ the space of all mixtures of Gaussian states, given an arbitrary mixed state $\rho$, we will find a Gaussian state $\sigma_{\rho} \in \mathcal{G}$ that is closest to $\rho$.
-As the entropy of $\rho$ increases, $\sigma_{\rho}$ will be get closer to $\rho$.
+Mathematically, the trace distance $T(\rho,\sigma)$ between two mixed states $\rho$ and $\sigma$ is contractive under a completely positive trace-preserving (CPTP) map $M$.
+That is, $T(M(\rho),M(\sigma)) \leq T(\rho,\sigma)$.\cite{nielsen2010quantum} In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive.
+In section 4, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite dimensional case, one could have infinite dimensional subspaces that remain orthogonal while increasing the entropy of all its elements.
+This corresponds, for example, to the case where there are conserved quantities.
+Note that in the map we construct in section 4, the expectation of all polynomials of position and momentum changes, and therefore no function of position and momentum is a conserved quantity.
+}  If we call $\mathcal{G}$ the space of all mixtures of Gaussian states, the states with a non-negative Wigner function, given an arbitrary mixed state $\rho$, we will find a state $\sigma_{\rho} \in \mathcal{G}$ that is closest to $\rho$.
+As we apply our entropy-increasing map, the entropy of $\rho$ increases and $\sigma_{\rho}$ will be get closer to $\rho$.
+In this sense the states will ``look classical.''
 
-What we described here is just a geometric property of the space of quantum mixed states, and it is independent on the source of entropy.
+What we described here is just a geometric property of the space of quantum mixed states, and it is independent of the source of entropy.
 
@@
 
-As mentioned, one of advantages of our approach is that it is compatible with more traditional approaches, including taking the limit $\hbar \to 0$.
-What one finds is that the math can typically be reinterpreted by taking an equivalent more physically meaningful limit.
+One of the advantages of our approach is that it is compatible with more traditional ones.
+Let us see how this works in two cases: the failure of classical statistical mechanics to predict black-body radiation and the recovery of classical statistical mechanics for thermal equilibrium.
 
-For example, following the original paper from Wigner \cite{WignerLimit}, we can consider the Wigner distribution of a system in thermal equilibrium at inverse temperature $\beta$:\footnote{TODO Man: there seem to be some difference from the original paper}
+\subsection{Black-body radiation}
+
+The failure of classical statistical mechanics to predict the black-body radiation spectrum was one of the key drivers for the development of quantum theory.
+The spectrum predicted by classical mechanics is given by the Rayleigh–Jeans law,
+\begin{equation}
+    \frac{2 \nu^2 k_\text{B} T}{c^2} = \frac{2 \nu^2}{c^2 \beta},
+\end{equation}
+where $\beta = \frac{1}{k_\text{B} T}$. The one predicted by quantum mechanics is given by Planck's law,
+\begin{equation}
+    \frac{ 2 h \nu^3}{c^2} \frac 1{\exp\left(\frac{h\nu}{k_\mathrm B T}\right) - 1} = \frac{ 2 h \nu^3}{c^2} \frac 1{e^{h \beta \nu} - 1}.
+\end{equation}
+It is well known that the two agree for small values of $\nu$ since $e^x \simeq 1 + x + O(x^2)$ for $x\ll 1$.
+\begin{equation}
+    \frac{ 2 h \nu^3}{c^2} \frac 1{e^{h \beta \nu} - 1} = \frac{ 2 h \nu^3}{c^2} \frac 1{1 + h \beta \nu + O\left(\nu^2\right) - 1} \simeq \frac{2 \nu^2}{c^2 \beta}.
+\end{equation}
+
+Note that the same result can be achieved taking the limit for large temperatures, i.e.~the classical black-body radiation spectrum can also be understood as the first term in the expansion for small $\beta$,
+\begin{equation}
+    \frac{ 2 h \nu^3}{c^2} \frac 1{e^{h \beta \nu} - 1} = \frac{ 2 h \nu^3}{c^2} \frac 1{1 + h \beta \nu + O\left(\beta^2\right) - 1} \simeq \frac{2 \nu^2}{c^2 \beta}.
+\end{equation}
+Since $\beta$ is the inverse of the temperature, this is the limit for large temperatures.
+Since the entropy increases as the temperature increases, this is also the limit for large entropy.
+That is, the classical distribution is recovered as the high-entropy limit for the quantum distribution.
+
+\subsection{Thermal equilibrium}
+
+We can similarly reinterpret results that take the mathematical limit $\hbar \to 0$.
+For example, following the original paper from Wigner \cite{WignerLimit}, we can consider the Wigner distribution of a system in thermal equilibrium at inverse temperature $\beta$:
 \begin{equation}
@@
 \end{equation}
-where $\epsilon(x,p)$ is the classical Hamiltonian.
-The extra term contains the quantum corrections; we will refer to it as $Q(\partial/\partial x)$.\footnote{TODO Man: Q vs Q hat} The Wigner function becomes
+where $\epsilon(x,p)$ is the classical Hamiltonian. The extra term contains the quantum corrections; we will refer to it as $\hat{Q}$. The Wigner function becomes
 \begin{equation}
@@
 We can show that quantum corrections are also only found at second order in $\beta$, justifying the classical limit.
-Let's start by considering the first order expansion, $e^{-\beta\tilde{H}}\simeq 1-\beta(\epsilon+\hat{Q})$.
+Let's start by considering the first-order expansion, $e^{-\beta\tilde{H}}\simeq 1-\beta(\epsilon+\hat{Q})$.
 So at first order
@@
 \end{equation}
-Focusing on the last term, we can insert a completeness (?) in the momentum eigenbase
+Focusing on the last term, we can insert an identity in the momentum eigenbase
 \begin{equation}
@@
 \end{equation}
-The first order quantum correction vanishes, showing that the classical approximation is correct to first order in $\beta$, which corresponds to the limit of large entropy.
-We can consider the second order correction
+The first-order quantum correction vanishes, showing that the classical approximation is correct to first order in $\beta$, which corresponds to the limit of large entropy.
+We can consider the second-order correction
 \begin{equation}
@@
 
-\section{The high entropy limit}
+\section{The high-entropy limit}
 
-We are now going to develop the limit in its most general formulation.
-TODO reword.
-Since the maximum entropy attainable by a quantum system is the logarithm of the dimension of the corresponding Hilbert space, the high entropy limit exists only for infinite-dimensional spaces.\footnote{TODO Man: add reference} This explains why spin, which lives in a finite-dimensional space, is an intrinsically quantum property.
+Having given qualitative reasons for the limit, and having seen how, in specific cases, standard arguments can be reinterpreted consistently with our view, we now develop the limit as a direct consequence of entropy increase.
+Since the maximum entropy attainable by a quantum system is the logarithm of the dimension of the corresponding Hilbert space,\cite{nielsen2010quantum} the high-entropy limit exists only for infinite-dimensional spaces.
+This explains why spin, which lives in a finite-dimensional space, is an intrinsically quantum property.
 Therefore we are going to concentrate on the case of a single degree of freedom identified by position and momentum.
-The case of multiple degrees of freedom can be recovered by increasing the entropy of all DOFs at the same time.
-That is, increasing the entropy of each DOF independently, without introducing correlations.
+The case of multiple degrees of freedom can be recovered by increasing the entropy of all DOFs at the same time, that is, by increasing the entropy of each DOF independently, without introducing correlations.
 
@@
 \begin{enumerate}
-    \item characterize entropy-increasing maps in classical mechanics for a single DOF, noting that they can all be understood as Hamiltonian evolutions followed by a pure stretching map, one that stretches position and momentum by the same factor $\sqrt{\lambda}$
-    \item show that a pure stretching map in quantum mechanics cannot be defined over symmetrized operator averages or normal ordering
-    \item show that it can be defined over anti-normal ordering; show that if we rescale the zero of entropy, the commutators become $[X, P]= \frac{\imath \hbar}{\lambda}$, which makes the limit $\lambda \to \infty$ mathematically equivalent to $\hbar \to 0$
-    \item show how, in the limit, the Wigner $W$ and Husimi $Q$ distributions become closer and closer
+    \item Characterize entropy-increasing maps in classical mechanics for a single DOF, noting that they can all be understood as Hamiltonian evolution followed by a pure stretching map, one that stretches position and momentum by the same factor $\sqrt{\lambda}$.
+Under this map, the expectation of a polynomial of position and momentum of degree $n$ is also multiplied by a factor of $\sqrt{\lambda^n}$.
+    \item Show that a pure stretching map in quantum mechanics cannot be defined over symmetrized operator averages or normal ordering.
+    \item Show that it can be defined over anti-normal ordering.
+Show that if we rescale the zero of entropy, the commutators become $[X, P]= \frac{\imath \hbar}{\lambda}$, which makes the limit $\lambda \to \infty$ mathematically equivalent to $\hbar \to 0$
+    \item Show how, in the limit, the Wigner $W$ and Husimi $Q$ distributions become closer and closer.
 \end{enumerate}
@@
 
-Before looking at the quantum case, let us study the high entropy limit of classical mechanics. We are looking for all those transformations that increase the entropy of all states by the same amount.
+Before looking at the quantum case, let us study the high-entropy limit of classical mechanics. We are looking for all those transformations that increase the entropy of all states by the same amount.
 
-We are going to study the one-dimensional case, therefore let us call $\mathcal{M} = (\mathbb{R}^2, \omega)$ the phase space for a single degree of freedom, where $\omega$ is the associated symplectic form.\footnote{The symplectic form allows us to define the count of states in infinitesimal regions, and therefore defines the entropy as well.} Suppose we have a map $R : \mathcal{M} \to \mathcal{M}$ that acts on phase space.
+We are going to study the one-dimensional case, therefore let us call $\mathcal{M} = (\mathbb{R}^2, \omega)$ the phase space for a single degree of freedom, where $\omega$ is the associated symplectic form.
+Suppose we have a map $R : \mathcal{M} \to \mathcal{M}$ that acts on phase space.
 This map will also act on probability distributions defined on $\mathcal{M}$ by moving them point to point.
-We require that $R$ increases the entropy of each distribution $\rho$ by a set value $\Delta S = \log \lambda > 0$.
-That is
+We require that $R$ increase the entropy of each distribution $\rho$ by a set value $\Delta S = \log \lambda > 0$.
+That is,
 \begin{equation}
@@
 The factor $\lambda$, then, can be understood either as the ratio between initial and final areas, or as the exponential of the entropy increase.
-Therefore studying the high entropy limit means studying what happens under stretching maps in the limit $\lambda \to \infty$.
+Therefore studying the high-entropy limit means studying what happens under stretching maps in the limit $\lambda \to \infty$.
 
@@
 For a pure stretching map, we can provide a more specific characterization.
-Note, in fact, that a distribution with compact support is always fully characterized by all its central moments, that is the expectations for all polynomials of position and momentum.
+Note, in fact, that a distribution with compact support is always fully characterized by all its central moments, that is, the expectations for all polynomials of position and momentum.
 Therefore, once we know how the central moments transform through the map, we know how all distributions with compact support transform and therefore we fully characterize the map.
@@
 Since, in general, $R(AB) \neq R(A)R(B)$, different permutations of polynomials will be transformed differently: the operator ordering matters.
-%GC: I think this is a simpler, more general argument. It also puts more in focus why I kept getting confused from the other argument
-
-% Note https://link.springer.com/article/10.1007/BF00419590 says entropy does not decrease if L L* \leq L* L. There also seems to be a bound
 
@@
 There are three commonly used orderings: symmetrized averages, normal ordering and anti-normal ordering.
-The first takes the product of $n$ operators by averaging all possible permutations; the second is in terms of polynomials of the form $(a^\dagger)^n a^m$, where $a=\sqrt{\frac{m\omega}{2\hbar}}(X+\frac{i}{m\omega}P)$ and $a^\dagger$ are the ladder operators; the last is in terms of $a^n (a^\dagger)^m$.
-The three orderings are associated, respectively, with a quasi-probability distribution: the Wigner function $W$, the Glauber-Sudarshan $P$ distribution and the Husimi $Q$ distribution.
+The first takes the product of $n$ operators by averaging all possible permutations; the second is in terms of polynomials of the form $(a^\dagger)^n a^m$, where $a=\sqrt{\frac{m\omega}{2\hbar}}(X+\frac{i}{m\omega}P)$ and $a^\dagger$ are the ladder operators, $m$ is the mass and $\omega$ the angular frequency; the last is in terms of $a^n (a^\dagger)^m$.
+The three orderings are associated, respectively, with a quasi-probability distribution:\cite{carmichael2013statistical} the Wigner function $W$, the Glauber-Sudarshan $P$ distribution and the Husimi $Q$ distribution.
 In each case, the expectation under the quasi-probability distribution of polynomials of classical variables returns the expectation of the respective ordering.
@@
 The normal ordering is ruled out as well.
-{\color{blue} thinking of it this way, you can also use the same argument as before.
-The P distribution also has negative regions that would stretch.} {\color{red} do we know they must have a maximum size?
-I haven't seen anything about it.}
 
@@
 \end{equation}
-where X is an operator, $L_i$ are the jump operators and $\gamma_i$ are positive parameters.
+where $X$ is an operator, $L_i$ are the jump operators and $\gamma_i$ are positive parameters.
 To define $T_Q$ for a single DOF we set $\gamma_i=\{\gamma\}$ and $L_i=\{a^\dagger\}$.
@@
 That is, while we fixed the ordering of the operators, the non-commutative nature of quantum observables still implies a choice within a one-parameter family of operators.
-However, note that a linear transformation that stretches $X$ and shrinks $P$
+However, note that a linear transformation that stretches $X$ and shrinks $P$ as
 \begin{equation}
@@
 \begin{equation}
-	\sum_i L_i.L_i^\dagger \leq \sum_i L_i^\dagger L_i.
+	\sum_i L_i L_i^\dagger \leq \sum_i L_i^\dagger L_i.
 \end{equation}
@@
     T_Q(X) &= T_Q\left(\sqrt{\frac{\hbar}{2m\omega}}(a^\dagger + a)\right) = \sqrt{\lambda} X \\
-    T_Q(P) &= T_Q(\imath\sqrt{\frac{\hbar m \omega}{2}}(a^\dagger - a)) = \sqrt{\lambda} P \\
-    [T_Q(X), T_Q(P)] &= \lambda [X, P].
+    T_Q(P) &= T_Q\left(\imath\sqrt{\frac{\hbar m \omega}{2}}(a^\dagger - a)\right) = \sqrt{\lambda} P \\
+    T_Q([X, P]) &= T_Q(\imath \hbar) = \imath \hbar \\
+    [T_Q(X), T_Q(P)] &= [ \sqrt{\lambda} X, \sqrt{\lambda} P] = \lambda [X, P] = \lambda \imath \hbar.
 \end{align}
-In other words, the effect of $T_Q$ is to increase the commutator between $X$ and $P$.
+Recall that the uncertainty principle between two operators is linked to the commutator by the formula
+\begin{equation}
+    \sigma_A \sigma_B = \frac{1}{2} | \langle [A,B] \rangle |.
+\end{equation}
+Therefore we find that the uncertainty between $T_Q(X)$ and $T_Q(P)$ grows with $\lambda$, which is consistent with the stretching behavior of the map.
 
-We can study the limit in two ways.
-The most straightforward would be see how the space changes as the commutators $[T_Q(X), T_Q(P)]= \lambda \imath \hbar$ go to infinity.
-This corresponds to keeping the entropy of the pure states fixed at zero and studying the space of states at high entropy.
-Alternatively, we can see how the space itself changes from the perspective of the state that is increasing in entropy.
-That is, we rescale the entropy of pure states to lower and lower values, while keeping the entropy of the mixed state fixed.
-To do that, we keep $[T_Q(X), T_Q(P)] = \imath \hbar$ constant and we study the limit as $[X, P]= \frac{\imath \hbar}{\lambda}$ goes to zero.
-Formally, this is equivalent to taking the limit $\hbar \to 0$, which is the group contraction that morphs the Moyal bracket Lie algebra to the Poisson bracket Lie algebra.\cite{saletan1961contraction, inonu1953contraction} That is, as $\lambda$ increases, the states will see the structure of quantum mechanics becoming closer and closer to the structure of classical mechanics.
+Note that when one takes the usual mathematical limit $\hbar \to 0$, the distribution, and therefore the uncertainty over position and momentum, is kept fixed.
+We can achieve that by redefining $X$ and $P$ in the following way
+\begin{align}
+    \hat{X} &= \frac{1}{\sqrt{\lambda}} X \\
+    \hat{P} &= \frac{1}{\sqrt{\lambda}} P \\
+    [\hat{X}, \hat{P}] &= \frac{\imath \hbar}{\lambda} \\
+    T_Q(\hat{X}) &=  X \\
+    T_Q(\hat{P}) &=  P \\
+    [T_Q(\hat{X}), T_Q(\hat{P})] &= \imath \hbar.
+\end{align}
+Note how the redefinition of the operators effectively undoes the stretching operation.
+Also note that the redefinition reduces the minimum uncertainty between position and momentum.
+The same limit, then, can be understood in two different ways.
+In the more physical viewpoint, we are mapping states within the same state space, which means pure states remain at the same level of uncertainty, of entropy.
+The map will move states to higher entropy within the same space as $\lambda$ increases.
+Conversely, in the more mathematical viewpoint, we can redefine the state space while going through the map.
+The uncertainty, the entropy, of the mapped states will remain the same, but the pure states of the target space will have lower  uncertainty, lower entropy, effectively adding new states at these lower values.
+As we take the limit $\lambda \to \infty$ in the first viewpoint, in the second viewpoint the commutator $[\hat{X},\hat{P}] = \frac{\imath \hbar}{\lambda}$ decreases, which is equivalent to taking the limit $\hbar \to 0$.
+This is the group contraction that morphs the Moyal bracket Lie algebra to the Poisson bracket Lie algebra.\cite{Moyal_1949,saletan1961contraction, inonu1953contraction} That is, as $\lambda$ increases, the entropy of pure states becomes lower and lower, and the structure of the quantum space becomes closer and closer to the structure of classical mechanics.
 
-%TODO: still missing a group contraction reference for \hbar \to 0
+\begin{figure}[h]
+    \centering
+\begin{tikzpicture}
+\begin{axis}[
+height=7cm,
+width=\linewidth*0.8,
+grid=both,
+grid style={line width=.1pt, draw=gray!25},
+axis lines=middle,
+xlabel = \(\Sigma\),
+ylabel = \(S\),
+legend style={at={(0.95,0.65)},anchor=east},
+]
+\addplot[thick,blue,samples=150,domain=0.01:1.5] {ln(x)+1};
+\addlegendentry{\(\text{classical}\)}
+\addplot[thick,BrickRed,densely dotted,samples=150,domain=0.5:1.5] {(x+1/2)*ln(x+1/2)-(x-1/2)*ln(x-1/2)};
+\addlegendentry{\(\text{quantum}\)}
+\addplot[thick,ForestGreen,dashed,samples=150,domain=0.25:1.5] {((2*x+1/2)*ln(2*x+1/2)-(2*x-1/2)*ln(2*x-1/2) - ln(2)};
+\addlegendentry{\(\text{$\lambda = 2$}\)}
+\addplot[thick,violet,densely dashed,samples=150,domain=0.05:1.5] {((10*x+1/2)*ln(10*x+1/2)-(10*x-1/2)*ln(10*x-1/2) - ln(10)};
+\addlegendentry{\(\text{$\lambda = 10$}\)}
+\end{axis}
+\end{tikzpicture}
+    \caption{Entropy $S$ in nats for a Gaussian state as a function of uncertainty $\Sigma$, measured in units of $\hbar$.
+Between the classical and the quantum case, we have the quantum case where pure states are rescaled by a factor of $\lambda$: $S_{\lambda Q}(\Sigma) = S_Q(\lambda \Sigma) - \ln(\lambda)$.
+Taking $\lambda$ to infinity corresponds to taking $\hbar$ to zero.}
+    \label{fig:uncertainty_scaled}
+\end{figure}
+
+As a more quantitative illustration, we can see what happens to the entropy of Gaussian states as we reduce the lowest possible uncertainty in the redefined quantum space.
+The entropy will become $S_{\lambda Q}(\Sigma) = S_Q(\lambda \Sigma) - \ln(\lambda)$.
+As we can see in figure \ref{fig:uncertainty_scaled}, even with low factors of $\lambda$ the entropy curve approaches the classical one.
+If one takes the limit $\lambda \to \infty$, the above expression reduces to the classical equation.
 
@@
 Also, while we have not shown that anti-normal ordering is the only ordering that allows a stretching map, it will need to have the same effect on $X$ and $P$, leading to the same limit.
-The question of whether all maps that perform the same group contraction can expressed as $T_Q \circ U$ remains open.
-However, they will factorize in that fashion in the limit, since the classical limit can always be factorized as pure stretching map following a symplectomorphism.
+The question of whether all maps that perform the same group contraction can be expressed as $T_Q \circ U$ remains open.
+However, they will factorize in that fashion in the limit, since the classical limit can always be factorized as a pure stretching map following a symplectomorphism.
 Similarly, the question of whether $T_Q$ increases the entropy of all states by the same amount remains open.
@@
 
-%Since the canonical operators x and p are simply linear combinations of $a$ and $a^\dagger$, their mean values will stretch as required, with stretching parameter given by $\lambda=e^{\gamma t}$.
-It is not so hard to show that the scaling applies to all operator combinations in anti-normal ordering.
-%If one is careful, it's clear that this cannot work forever.
-We can stretch the first moment of the operators: $a_i^{(\dagger)}(t)=\sqrt{\lambda} a_i^{(\dagger)}(0)$.
-But already for the second moment this breaks.
-We find: $(a^\dagger a)(t)=\lambda((a^\dagger a)(0)+1)-1$.
-The exact stretching is only recovered in the limit of  $\lambda\rightarrow\infty$.
-On the other hand, if we consider the anti-normal ordering $(a a^\dagger)(t)=\lambda(a a^\dagger)(0)$.
-This means that the map we have satisfies stretching in anti-normal ordering.
-
-%The correct distribution function to use under anti-normal ordering is the Husimi Q-distribution.
-This makes perfect sense.
-This is the only one that is positive-defined.
-Therefore it's the only one that can be stretched without issues due to negative regions
-
-%TODO: show that such a map exist and it increases the entropy by a fixed amount (DONE?)
-
-% Possible resource https://link.springer.com/article/10.1007/BF00419590 to show map increase entropy
-
 \subsection{Stretching the Wigner distribution}
 
-To understand what happens physically during the group contraction, let us see how the Wigner W distribution changes under a pure quantum stretching map.
-To do that, we can use the fact that Q is the Weierstrass transform of $W$.
+To understand what happens physically during the group contraction, let us see how the Wigner $W$ distribution changes under a pure quantum stretching map.
+To do that, we can use the fact that $Q$ is the Weierstrass transform of $W$.
 That is:
@@
     F\left(\frac{1}{\lambda}f\left(\frac{\alpha}{\sqrt{\lambda}}\right)\right)(k)=F(f(\alpha))(\sqrt{\lambda}k)
-\end{equation} and using the convolution theorem, we get (TODO Man: check)
+\end{equation} and using the convolution theorem, we get
 \begin{equation}
@@
 \end{equation}
-We now notice that (TODO Man: add (k) in the end?)
+We now notice that
 \begin{equation}
-    \frac{F(G)(\sqrt{\lambda}k)}{F(G)(k)}=\frac{e^{-\lambda|k|^2/8}}{e^{-|k|^2/8}}=e^{-(\lambda-1)|k|^2/8}=F(G)(\sqrt{\lambda-1}k)=F\left(\frac{1}{\lambda-1}G\left(\frac{\beta}{\sqrt{\lambda-1}}\right)\right),
+    \frac{F(G)(\sqrt{\lambda}k)}{F(G)(k)}=\frac{e^{-\lambda|k|^2/8}}{e^{-|k|^2/8}}=e^{-(\lambda-1)|k|^2/8}=F(G)(\sqrt{\lambda-1}k)=F\left(\frac{1}{\lambda-1}G\left(\frac{\beta}{\sqrt{\lambda-1}}\right)\right)(k),
 \end{equation}
@@
 A clear interpretation can be made by working directly on the Fourier transform of $W$; see Eq.~(\ref{Fourier}).
-The function $F(W_\lambda)$ is a scaled version of $F(W_1)$, with a Gaussian filter applied to it.
+The function $F(W_\lambda)$ is a scaled version of $F(W_1)$ with a Gaussian filter applied to it.
 Quantum information in $W_1$ is carried by spectral weights with k-vectors larger than 1.
@@
 
-We have seen that classical mechanics can be recovered as the high entropy limit of quantum mechanics.
-That is, states of high entropy are better and better approximated by classical distributions over phase space.
+We have seen that classical mechanics can be recovered as the high-entropy limit of quantum mechanics.
+That is, states of higher and higher entropy are better and better approximated by classical distributions over phase space.
 This approach to the classical limit is independent of mechanism and interpretation, as it does not matter how the entropy is increased or what one believes quantum states to represent: as long as the description is in terms of mixed states of sufficiently high entropy, classical mechanics applies.
@@
 Physically, this can be understood as taking the entropy of pure states to minus infinity, which is equivalent to saying that the relative entropy of mixed states goes to plus infinity.
+That is, in the same way that $c \to \infty$ should be understood as $v \ll c$, $\hbar \to 0$ should be understood as $S \gg 0$, as physically we cannot take limits of physical constants.
 This gives a reasonable and precise account of the classical limit: in the same way that non-relativistic mechanics applies to low speed, classical mechanics applies to high entropy.
@@
 
-In this light, we wonder whether it can be proven that quantum mechanics is the only way to fix the low entropy range of classical mechanics.
+In this light, we wonder whether it can be proven that quantum mechanics is the only way to fix the low-entropy range of classical mechanics.
 That is, is quantum mechanics the only theory that can recover classical mechanics at high entropy?

### 2024-11-05 Articles/2023-HilbertQuantum/Submissions/Quantum Studies/HilbertQuantum.tex
HilbertQuantum: better discussion of rigged HS

@@
 
-To make this determination, we need to provide a standard to judge the physicality of a formal definition.\cite{redei_tension_2020, north_physics_2021} Suppose we are studying a physical system.
+To make this determination, one needs a standard to judge the physicality of a formal definition\cite{redei_tension_2020, north_physics_2021} and, rather tellingly, physics tradition does not provide one.
+Here is the standard we follow in our work.\cite{aop-book}  Suppose we are studying a physical system.
 A mathematical definition is \textbf{physical} if it properly characterizes the physical system in the given conditions.
@@
 This is the task at hand.
-
-Some mathematicians were disappointed that we didn't discuss in depth approaches like rigged Hilbert spaces or the algebraic approach to quantum mechanics.
-Do rigged Hilbert spaces spell out exactly what each element of the Gelfand triple describes and give a physical justification that tells us exactly which of those are the physically correct specifications?
-Do algebraic approaches give a physical justification as to why things we measure in a lab must be modeled by a $C^*$-algebra, and consequently allow us to determine whether the use of Hilbert spaces to model states can be given (or not given) a physical justification?
-To our knowledge, and to the knowledge of the mathematicians that approached us, they do not.
-This does not mean that these works are of no interest to a physical audience.
-It just means they are not going to help with the task at hand.
 
@@
 
-Obviously, while the argument is ultimately physical, and it is therefore the job of a physicist to make these arguments, it has to be a physicist that is well-versed in foundational issues in mathematics.
+If the problem of justifying the use of mathematical structures in physics is, ultimately, a physical problem, it also follows that it cannot be solved simply by a different mathematical approach.
+For example, according to the literature, (CITE) rigged Hilbert spaces were developed to give a mathematical foundation to Dirac’s bra-ket formalism, and be able to talk about eigenstates of position and momentum.
+The question is not whether these can be useful at a mathematical level to solve certain problem.
+They clearly are.
+The question is whether, on physical grounds, this is a good idea.
+Should we really treat continuous and discrete spectra the same, or realize that they are, on physical grounds, two rather different things?
+Should the fact that the energy spectrum is typically discrete for bound states and continuous for unbound states more seriously, and realize that there is something in the topology of the spectra that is rather physically important?
+Should we then make the different more prominent in our thinking, and not hide in mathematical details?
+Similarly, one can take the algebraic route and start with an algebra of observables.
+This, some argue, make some mathematical problems easier to solve.
+Again, no question about it.
+It still begs the question: why is it that, on physical grounds, observables should form a $C^*$-algebra?
+What, then, is the suitable physically meaningful state space?
+Moreover, note that in both those approaches Hilbert spaces still play a crucial role.
+If Hilbert spaces are physically problematic, these approaches are still relying on an object that is physically problematic.
+
+While the argument is ultimately physical, and it is therefore the job of a physicist to make these arguments, it has to be a physicist that is well-versed in foundational issues in mathematics.
 The justification, in fact, will need precise understanding of the mathematical definitions and their implication for both the physics and the math.
+It is often the case that the mathematical detail informs the physics and shouldn't be swept under the rug.
 Let us give an example.
@@
 The fact that other tools have been developed to solve other mathematical problems is not relevant to this discussion.
-We are only interested in the question of mathematical representation of quantum states, \textbf{physical} entities we prepare in a lab.
+We are only interested in the question of mathematical representation of quantum states, which are physical entities we prepare in a lab.
 Something we produce, manipulate and, in the end, measure.

### 2024-11-06 Articles/2023-HilbertQuantum/Submissions/Quantum Studies/HilbertQuantum.tex
HilbertQuantum: describing what states are
Further edits in response to reviewer comments

@@
 When we first drafted this paper, we naively thought that the advantage of having a well-posed mathematical structure that represents, and only represents, physically well-defined objects would be self-evident.
-It does not seem to be the case: a previous reviewer stated that ``the question of whether the limiting states are actually physical is rather pointless;'' conversely, others see no problem in working with QFT path integrals that are not mathematically well-defined.
-We are therefore in the absurd, from our perspective, position of having to defend that, yes, the math we need to use in physics has be mathematically precise and, yes, the math we need to use in physics has to represent, perhaps idealized, physically realizable objects.
-The cost of this failure is all around us: an ever increasing number of interpretations of quantum mechanics, abstract mathematical work whose connection to physics is unclear, theoretical physics more and more disconnected from experimental physics.
-The reader that, like us, finds this obvious can skip this section.
+It does not seem to be the case: a reviewer from a previous journal submission stated that ``the question of whether the limiting states are actually physical is rather pointless;'' conversely, others see no problem in working with quantum field theory path integrals that are not mathematically well-defined.
+We are therefore in the absurd, from our perspective, position of having to defend that, yes, the math we need to use in physics has be mathematically precise and, yes, the math we need to use in physics has to represent---perhaps idealized---physically realizable objects.
+The cost of this failure is all around us: an ever-increasing number of interpretations of quantum mechanics, abstract mathematical work whose connection to physics is unclear, theoretical physics more and more disconnected from experimental physics.
+The reader who, like us, finds this obvious can skip this section.
 
-One problem is that most physicists see mathematics as just a tool for calculation, much like spectroscopes or telescopes are tools to gather data.
+One problem is that most physicists see mathematics as just a tool for calculation, much like oscilloscopes or telescopes are tools to gather data.
 As long as they do what they are supposed to do, the details are irrelevant because the physics lies somewhere else.
@@
 But mathematics plays a much more fundamental role in physics: it is the language in which we specify our models.
-When we say ``states are rays in a Hilbert space'' we are giving a physical specification.
+When we say, ``States are rays in a Hilbert space,'' we are giving a physical specification.
 We are saying that there are objects in the real world that can be modeled with a particular mathematical structure.
@@
 To make this determination, one needs a standard to judge the physicality of a formal definition\cite{redei_tension_2020, north_physics_2021} and, rather tellingly, physics tradition does not provide one.
-Here is the standard we follow in our work.\cite{aop-book}  Suppose we are studying a physical system.
+Here we describe the standard that we follow in our work.\cite{aop-book}  Suppose we are studying a physical system.
 A mathematical definition is \textbf{physical} if it properly characterizes the physical system in the given conditions.
-That is, if we are able to \textbf{justify on physical grounds} that the given mathematical definition is needed to capture and only capture a particular aspect of the physical system under the given condition.
+That is, if we are able to \textbf{justify on physical grounds} that the given mathematical definition is needed to capture and only capture a particular aspect of the physical system under the given conditions.
 On the other hand, a mathematical definition is \textbf{unphysical} if it can be shown to require properties or operations that cannot have a physical counterpart.
@@
 In our specific case, the core argument we will use is that states described by measurable quantities with infinite or ill-defined expectation values are not physically realizable.
-To us, and to every experimental physicist we talked to, ``it doesn't make sense experimentally,'' and this is the best argument possible to show that something is unphysical.
+To us, and to every experimental physicist we talked to, ``It doesn't make sense experimentally,'' and this is the best argument possible to show that something is unphysical.
 ``It can't be done.'' No argument is better.
-However, more than one previous reviewer found no problem with infinite (or undefined) expectation values, as they adopt an ``ensemble interpretation,'' and tried to convince us that we shouldn't either.
-How can progress on the mathematical foundations of quantum mechanics be made if an argument that is accepted by the vast majority of physicists is rejected?
+However, more than one previous reviewer found no problem with infinite (or undefined) expectation values, as they adopt an ``ensemble interpretation,'' and tried to convince us that we should not be concerned either.
+How can progress on the mathematical foundations of quantum mechanics be made if an argument that is accepted by the vast majority of physicists is rejected when one tries to discuss it in the literature?
 
 If the problem of justifying the use of mathematical structures in physics is, ultimately, a physical problem, it also follows that it cannot be solved simply by a different mathematical approach.
-For example, according to the literature, (CITE) rigged Hilbert spaces were developed to give a mathematical foundation to Dirac’s bra-ket formalism, and be able to talk about eigenstates of position and momentum.
-The question is not whether these can be useful at a mathematical level to solve certain problem.
+For example, rigged Hilbert spaces were developed to give a mathematical foundation to Dirac’s bra-ket formalism\cite{Dirac1947}, and to be able to talk about eigenstates of position and momentum, working mathematically with both continuous and discrete bases on an interchangeable level\cite{Madrid_2005,Celeghini2016}.
+The question is not whether these can be useful at a mathematical level to solve certain problems.
 They clearly are.
-The question is whether, on physical grounds, this is a good idea.
-Should we really treat continuous and discrete spectra the same, or realize that they are, on physical grounds, two rather different things?
-Should the fact that the energy spectrum is typically discrete for bound states and continuous for unbound states more seriously, and realize that there is something in the topology of the spectra that is rather physically important?
-Should we then make the different more prominent in our thinking, and not hide in mathematical details?
+The question at hand is whether, on physical grounds, this is a good idea.
+Should we really treat continuous and discrete spectra the same, or should we consider that they are, on physical grounds, two rather different things?
+Should the fact that the energy spectrum is typically discrete for bound states and continuous for unbound states make us wonder if there is something in the differing topologies of the spectra that is physically important?
+Should we then make the difference more prominent in our thinking, rather than trying to construct mathematical tools that treat them the same?
 Similarly, one can take the algebraic route and start with an algebra of observables.
-This, some argue, make some mathematical problems easier to solve.
-Again, no question about it.
+This, some argue, makes some mathematical problems easier to solve, which indeed it does.
 It still begs the question: why is it that, on physical grounds, observables should form a $C^*$-algebra?
 What, then, is the suitable physically meaningful state space?
-Moreover, note that in both those approaches Hilbert spaces still play a crucial role.
-If Hilbert spaces are physically problematic, these approaches are still relying on an object that is physically problematic.
+Moreover, note that in both these frameworks, Hilbert spaces still play a crucial role.
+If Hilbert spaces are physically problematic, these approaches are still relying on a mathematical structure that is physically problematic.
 
 While the argument is ultimately physical, and it is therefore the job of a physicist to make these arguments, it has to be a physicist that is well-versed in foundational issues in mathematics.
-The justification, in fact, will need precise understanding of the mathematical definitions and their implication for both the physics and the math.
-It is often the case that the mathematical detail informs the physics and shouldn't be swept under the rug.
+The justification, in fact, will need precise understanding of the mathematical definitions and their implications for both the physics and the math.
+It is often the case that the mathematical detail informs the physics and should not be swept under the rug.
 Let us give an example.
-In a previous work we found that a set of physically distinguishable objects must be a $T_0$ second countable topological space where each open set represents a verifiable statement, a statement allowing a test that terminates in finite time if and only if the statement is true.
-This result tells us why functions in physics must be ``well-behaved'' (i.e.~topologically continuous in the right topology), why sets with cardinality greater than the continuum are unphysical (i.e.~they cannot be given a $T_0$ second countable topology) and why the Banach-Tarski paradox does not apply in physics (i.e.
-non-Borel sets are unphysical as they cannot be associated with an experimental test).
+In a previous work we found that a set of physically distinguishable objects must be a $T_0$ second-countable topological space where each open set represents a verifiable statement, a statement allowing a test that terminates in finite time if and only if the statement is true\cite{aop-book}.
+This result tells us why functions in physics must be ``well-behaved'' (i.e.~topologically continuous in the right topology), why sets with cardinality greater than the continuum are unphysical (i.e.~they cannot be given a $T_0$ second-countable topology) and why the Banach-Tarski paradox does not apply in physics (i.e.~non-Borel sets are unphysical as they cannot be associated with an experimental test).
 This tells us that, at the very least, non-separable Hilbert spaces are unphysical as they have cardinality greater than the continuum.\cite{pittphilsci18363} This ties together ideas from point-set topology, logic, set theory, measure theory and computer science.
-Clearly, one needs to have at least a basic understanding of these fields to be able to follow the argument, and most degree programs in physics do not include typically discuss these topics in depth.\footnote{A previous reviewer conflated the completeness of the Hilbert space with the existence of a complete orthonormal basis of said space.} However, in the end, only a physicist can judge whether the original premise, the evidence-based requirement of physics, justifies the mathematical requirements.
+Clearly, one needs to have at least a basic understanding of these fields to be able to follow the argument, and most degree programs in physics do not typically discuss these topics in depth.\footnote{A previous reviewer conflated the completeness of the Hilbert space with the existence of a complete orthonormal basis of said space.} However, in the end, only a physicist can judge whether the evidence-based requirement of physics justifies the mathematical requirements.
 
-To summarize, we are not interested in understanding whether Hilbert spaces have nice mathematical features for calculating or proving results.
-The fact that other tools have been developed to solve other mathematical problems is not relevant to this discussion.
+To summarize, we are not interested here in understanding whether Hilbert spaces have nice mathematical features for calculating or proving results.
+Moreover, the fact that other tools have been developed to solve other mathematical problems is not relevant to this discussion.
 We are only interested in the question of mathematical representation of quantum states, which are physical entities we prepare in a lab.
-Something we produce, manipulate and, in the end, measure.
+States in physics are the output of a preparation procedure, a characterization that works for both ``pure states'' of mechanics as well as ensembles in statistical mechanics.
+They are what we produce, manipulate and, in the end, measure.
 What physical requirements are we capturing when we model them mathematically?

### 2024-11-07 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed intro

@@
 \usepackage{breakurl}
+\usepackage{tcolorbox}
 
@@
 
-As mentioned in the introduction, our goal is to construct a probability theory that is guaranteed to work in any physical theory.
-Therefore our fundamental axioms will not be about probability, but about physical theories.
-This will help us place probability in a bigger context, and provide a minimal interpretation.
-As it is common in our larger project, we want axioms to describe necessary requirements of the scientific practice.
-That is, they will describe features of physical practice, not feature of the systems being described.
+As mentioned in the introduction, our goal is to construct a probability theory of general applicability to physical systems.
+Our approach is not to define an axiomatic theory of probability and apply to physical systems, but rather characterize basic properties of physical states and from those build a theory of probability that fits those definitions.
+As it is common in our larger project, we start from axioms that, in our opinion, represent requirements of the scientific practice.
+The advantage is that the realm of applicability and physical interpretation will be clear from the start.
 
-We are looking to find a minimal set of axioms required by a physical theory that allows statistical descriptions.
-Such a theory must, at the very least, provide us with a notion of ensemble, and the space of ensembles that are allowed in the theory.
+We start from requiring a physical theory to allow statistical descriptions: at the very least, it will provide us with a notion of ensemble, and the space of ensembles that are allowed by the theory.
 An ensemble can be understood as the infinite collection of the outputs of a particular preparation procedure.
@@
 In quantum mechanics, instead, it is the space of density operators.
-While the ensemble space for each theory will be in principle different, some will have many common features.
-In all cases we must be able to talk about conservative or dissipative processes, measurable properties of the system, probability distributions for outcomes of said properties and so on.
+While the ensemble space for each theory will be in principle different, they will all share common features: they must allow measurable properties of the system, probability distributions for outcomes of said properties, characterize reversible and irreversible processes through an entropy function, and so on.
 We are looking for necessary minimal basic requirements upon which these common tools can be built.
 
-At least two arguments can be made to show that any physical theory must provide a set of ensembles.
+At least two arguments can be made to show that any physical theory must allow statistical descriptions, provide an ensemble space.
 First, note that physical laws are never about specific instances but rather regularities.
 They are statements of the type ``whenever we prepare this, we can measure that.'' This means that, ultimately, they are about ensembles.
-Second, if a physical theory is to be testable experimentally, and repeatedly so, it must be in terms of ensembles because this is what we prepare in practice and this is what, in the end, we characterized experimentally.
-It seems, then, that the approach can be given firm conceptual grounds, which we leave as open but solvable philosophical problem.
+Second, if a physical theory is to be testable experimentally, and repeatedly so, it must be in terms of ensembles because this is what we prepare in practice and this is what, in the end, we characterize experimentally.
+The approach, then, can most likely be given firm conceptual grounds, which we leave as open but solvable philosophical problem.
 
@@
 
+In the case of ensembles, typical verifiable statements will be ``the average energy of the particle is between 3 and 4 $eV$,'' ``the probability of getting heads is between 49 and 51 percent,'' ``the probability distribution for the position is a Gaussian of average 0 $m$ and 1 $m$ standard deviation with 1\%.'' Note that these are not results of single shot measurements, and therefore verifiable statements are distinctly different from quantum measurements.
+In the same manner, these are not measurements that extract one bit of information, as there is no requirement of termination in the negative case.
+
+\begin{tcolorbox}[colback=white, colframe=black]
+	Experimental verifiability $\Rightarrow$ An ensemble space $\Ens$ is a $T_0$ second-countable topological space.
+\end{tcolorbox}
+
 \subsection{Statistical mixtures}
 Another basic requirement is the ability to prepare a \textbf{mixture} of two ensembles.
-If $\ens[a], \ens[b] \in \Ens$ are two ensembles, and $p \in [0,1]$ a weight, then $\ens = p \ens[a] + (1-p) \ens[b] \in \Ens$ that describes a process that selects the first ensemble over the second $p$ percent of the times.
+If $\ens[a], \ens[b] \in \Ens$ are two ensembles, and $p \in [0,1]$ a weight, then $\ens = p \ens[a] + (1-p) \ens[b] \in \Ens$ is the ensemble that describes a process that selects the first ensemble over the second $p$ percent of the times.
 Mathematically, the ensemble space is endowed with a \textbf{convex structure}.
+Moreover, since the mixing operation must be consistent with experimental verifiability, it will be topologically continuous an the ensemble space will be a topological convex space.
 
@@
 
+\begin{tcolorbox}[colback=white, colframe=black]
+	Statistical mixture of ensemble $\Rightarrow$ An ensemble space $\Ens$ is a topological convex space.
+\end{tcolorbox}
+
 \subsection{Entropy}
-The last requirement is that each ensemble must have a well defined entropy. That is, there is a scalar function $S : \Ens \to \mathbb{R}$ that satisfies the following
+The last requirement is that each ensemble must have a well defined \textbf{entropy}. That is, there is a scalar function $S : \Ens \to \mathbb{R}$ that satisfies the following
 \begin{enumerate}
@@
 The first tells us that, during mixing, the entropy cannot decrease, and it stays the same if and only we are mixing an ensemble with itself.
-The second tells us that the most the entropy can increase is given by the choice between the two ensembles.\footnote{The use of the Shannon entropy for the upper bound can likely be recovered.}
+The second tells us that the most the entropy can increase is given by the choice between the two ensembles.\footnote{It can be shown that if an upper bound that only depends on the mixing coefficients exists, then it must be the Shannon entropy.}
 
 The entropy is ultimately responsible for all geometrical structure in physics.
-In classical mechanics, the geometry is essentially defined by volumes in phase space.
+In classical mechanics, the geometry is essentially defined by phase space volumes and areas in each DOF (degree of freedom).
 The entropy of uniform distributions is given by the logarithm of the volume, meaning that given the volumes we are able to calculate the entropy and given the entropy we are able to reconstruct the volume.
+Similarly, uniform marginals will recover DOF areas.
 In quantum mechanics, the inner product (i.e.
-the Born rule) allows us to calculate the entropy of the mixture of two pure states, and given that entropy we can recover the Born rule.
+the Born rule) allows us to calculate the entropy of the mixture of two pure states, and given the entropy we can recover the Born rule.
 
 An interesting insight is that the entropy, being strictly concave, has a negative defined Hessian.
-The negation of the Hessian, then, is a symmetric positive defined function of two infinitesimal variations, and can serve as a metric tensor.
+The negation of the Hessian, then, is a symmetric positive defined function of two infinitesimal variations, and can serve as a metric tensor on the affine structure given by the mixing coefficients.
 In both classical and quantum mechanics this recovers the Fisher-Rao metric.
@@
 Therefore we define two ensembles to be \textbf{orthogonal} if their mixture lead to a maximal entropy increase.
-If two states are orthogonal, on physical grounds, the must have no common component.
-However, the reverse is not case: two pure states in quantum mechanics have no components to have in common, but are not necessarily orthogonal.
-However, classical probability spaces are exactly those space in which no common component implies orthogonality.
-This is the main source of the ``weirdness'' of quantum mechanics.
+An additional axiom is valid for orthogonal ensembles.
+We say an ensemble is a component of another ensemble if the second can be expressed as a mixture of the first with some other ensemble.
+On physical grounds, two orthogonal ensembles $\ens_1, \ens_2 \in \Ens$ are perfectly distinguishable and must have no common component.
+That is, there is no $\ens_3 \in \Ens$ that is a component of both $\ens_1$ and $\ens_2$.
+Note that the converse is not generally true: two pure states in quantum mechanics have no components to have in common, but are not necessarily orthogonal.
+We find that an ensemble space where no common component implies orthogonality can be described by a space of classical probability measures.
+Non-classicality, then, is exactly the failure of this property.
+
+\begin{tcolorbox}[colback=white, colframe=black]
+	Existence of entropy $\Rightarrow$ Strictly concave function whose upper bound defines orthogonality.
+\end{tcolorbox}
 
@@
 Mathematically, it would be very convenient.
-However, not all infinite mixtures correspond to physically meaningful systems, such as a particle with an infinite average position.
+However, some infinite mixtures lead to physically ill-defined systems, such as a particle with an infinite average position.
 It is the connection to experimental verification, the topology, that will tell us which limits are allowed and which are not.
-This type of conceptual clarity has to come first.
-The math follows.
+Conceptual clarity on these details is essential to find the correct generalization.
 A full examination of these issues is beyond the scope of this article, and we refer to our larger, and future, work.
@@
 For example, the pure states for a six-sided die are the ensemble perfectly prepared respectively on the six possible outcomes.
-All ensembles are convex combinations $\sum_{i+1}^{6} p_i = 1$ of those pure states.
+All ensembles are convex combinations $\sum_{i=1}^{6} p_i \ens_i = 1$ of those pure states $\ens_i$.
 Mathematically, the pure states are the extreme points of the space and all other points can be understood as convex combinations over them.
-This is exactly how classical discrete spaces work.
-Most physicists, including are older selves, think that that's how everything works.
-For example, in classical mechanics the pure states are perfect preparations of position and momentum, and ensembles are probability distributions.
-This does not work.
+This is exactly how classical discrete spaces work and there is a temptation to think this is how everything works.
+For example, in classical mechanics the pure states would be perfect preparations of position and momentum, and ensembles are probability distributions.
+Not only this is problematic mathematically, it is conceptually not sustainable.
 
-We cannot consistently prepare classical particles perfectly, therefore the points of phase space are not ensemble.
-The mathematical facts that delta Dirac are not proper functions, have support with empty interior, have minus infinite entropy, and so on, are simply a consequence of our conceptual problem.
-A bit better is to think them as the limit of infinitesimal decomposition: classical ensembles can be decomposed into more and more precise ones, and the points of are the limit of this decomposition.
+On physical grounds, we cannot consistently prepare classical particles perfectly, therefore there is no ensemble corresponding to a perfect preparation of a point of phase space.
+The mathematical facts that delta Dirac are not proper functions, their support has an empty interior, have minus infinite entropy, and so on, are simply a consequence of our conceptual problem.
+One may concede that the ensembles to do not exist physically, but may still want them to consider them as the limit of infinitesimal decomposition: classical ensembles can be decomposed into more and more precise ones, and the points of are the limit of this decomposition.
 Probability distributions are defined on these limits.
-This works in classical mechanics, but in general.
+This kind of works in classical mechanics, but in general.
 
-In quantum mechanics we have pure states, and these are ensembles that cannot be further decomposed into finer ensembles.
+In quantum mechanics we do have pure states, and these are ensembles that cannot be further decomposed into finer ensembles.
 Therefore an eigenstate of position is not a limit of decomposition, a limit in the convex structure.
@@
 It is the limit of a sequence of pure states, a limit in the topology.
-Physically, more than mathematically, it is a completely different thing, therefore it is not a good route to generalization.\footnote{The fact that pure states can be seen, in some sense, as superposition of eigenstates of position, makes the matter even more confusing.}
+Physically, more than mathematically, it is a completely different thing.
+It is only when we realized this difference that a clean route to generalization presented itself.\footnote{The fact that pure states can be seen, in some sense, as superposition of eigenstates of position, makes the matter even more confusing.}
 
@@
 open) interval of the possible values of said quantity (its spectra).
-The two coincide in classical discrete space, which is the least physically interesting case as it does not even allow for classical particle mechanics.
+The two coincide in classical discrete spaces, which are the least physically interesting case as they do not even allow classical particle mechanics.
+Equivalently, \textbf{mixtures are about preparations, not measurements, and probability distributions are about measurements, not preparations}.
+Since they can only coincide in classical spaces, a generalized notion of probability for physical theories must resolve and characterize the distinction.
 

### 2024-11-09 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: typos

@@
 
-In the case of ensembles, typical verifiable statements will be ``the average energy of the particle is between 3 and 4 $eV$,'' ``the probability of getting heads is between 49 and 51 percent,'' ``the probability distribution for the position is a Gaussian of average 0 $m$ and 1 $m$ standard deviation with 1\%.'' Note that these are not results of single shot measurements, and therefore verifiable statements are distinctly different from quantum measurements.
+In the case of ensembles, typical verifiable statements will be ``the average energy of the particle is between 3 and 4 $eV$,'' ``the probability of getting heads is between 49 and 51 percent,'' ``the probability distribution for the position is a Gaussian of average 0 $m$ and 1 $m$ standard deviation within 1\%.'' Note that these are not results of single shot measurements, and therefore verifiable statements are distinctly different from observables in quantum mechanics.
 In the same manner, these are not measurements that extract one bit of information, as there is no requirement of termination in the negative case.
@@
 Mathematically, the ensemble space is endowed with a \textbf{convex structure}.
-Moreover, since the mixing operation must be consistent with experimental verifiability, it will be topologically continuous an the ensemble space will be a topological convex space.
+Moreover, since the mixing operation must be consistent with experimental verifiability, it will be topologically continuous and the ensemble space will be a topological convex space.
 
@@
 If we fix $\ens$, $\ens[a]$ and $p$ then, if it exists, there is only one $\ens[b]$ such that $\ens = p \ens[a] + (1-p) \ens[b]$.
-The ensemble space is \textbf{complemented}, meaning that there is only one complement to a component of an ensemble.
+The ensemble space is \textbf{complemented}, meaning that there is only one complement to a component of an ensemble.\footnote{In the context of convex spaces, this property is called ``cancellative.''}
 
-This perspective allows to understand what negative probability is.
+This perspective allows us to understand what negative probability is.
 Suppose we have a real vector space of dimensions $n$, then $n$ linearly independent vectors are enough to express every point as a linear combination.
 If we add another point, we can now express every point as an affine combination, that is a linear combination where the coefficients sum to one, though some can still be negative.
-Since the space of mixed space in quantum mechanics is a real vector space, we can choose a linearly dependent set of states space span the whole space, and every mixed state can be expressed as an affine combination, as a pseudo-probability.
+Since the space of mixed states in quantum mechanics is a real vector space, we can choose a linearly dependent set of states that span the whole space, and every mixed state can be expressed as an affine combination, as a pseudo-probability.
 
@@
 \begin{enumerate}
-	\item strictly convex: $S(p \ens[a] + (1-p) \ens[b]) - (p S(\ens[a]) + (1-p) S(\ens[b]) ) \geq 0$
-	\item upper bounded: $S(p \ens[a] + (1-p) \ens[b]) - (p S(\ens[a]) + (1-p) S(\ens[b]) ) \leq - p \log p - (1-p) \log(1-p)$
+	\item strictly concave: $S(p \ens[a] + (1-p) \ens[b]) - (p S(\ens[a]) + (1-p) S(\ens[b]) ) \geq 0$
+	\item bounded from above: $S(p \ens[a] + (1-p) \ens[b]) - (p S(\ens[a]) + (1-p) S(\ens[b]) ) \leq - p \log p - (1-p) \log(1-p)$
 \end{enumerate}
@@
 Another important feature is that if two ensembles saturate the upper bound, then they are orthogonal in both classical and quantum mechanics.
-Therefore we define two ensembles to be \textbf{orthogonal} if their mixture lead to a maximal entropy increase.
+Therefore we define two ensembles to be \textbf{orthogonal} if their mixture leads to a maximal entropy increase.
 An additional axiom is valid for orthogonal ensembles.
@@
 That is, there is no $\ens_3 \in \Ens$ that is a component of both $\ens_1$ and $\ens_2$.
-Note that the converse is not generally true: two pure states in quantum mechanics have no components to have in common, but are not necessarily orthogonal.
+Note that the converse is not generally true: two pure states in quantum mechanics have no components in common, but are not necessarily orthogonal.
 We find that an ensemble space where no common component implies orthogonality can be described by a space of classical probability measures.

### 2024-11-09 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Fix some typos before copying parts to NSF proposal

@@
 As mentioned in the introduction, our goal is to construct a probability theory of general applicability to physical systems.
-Our approach is not to define an axiomatic theory of probability and apply to physical systems, but rather characterize basic properties of physical states and from those build a theory of probability that fits those definitions.
-As it is common in our larger project, we start from axioms that, in our opinion, represent requirements of the scientific practice.
+Our approach is not to define an axiomatic theory of probability and apply it to physical systems, but rather to characterize basic properties of physical states and from those build a theory of probability that fits those definitions.
+As is common in our larger project, we start from axioms that, in our opinion, represent requirements of scientific practice.
 The advantage is that the realm of applicability and physical interpretation will be clear from the start.
@@
 An ensemble can be understood as the infinite collection of the outputs of a particular preparation procedure.
-The ensemble space of classical mechanics, for example, is the space of continuous probability distributions of phase space.
+The ensemble space of classical mechanics, for example, is the space of continuous probability distributions on phase space.
 In quantum mechanics, instead, it is the space of density operators.
-While the ensemble space for each theory will be in principle different, they will all share common features: they must allow measurable properties of the system, probability distributions for outcomes of said properties, characterize reversible and irreversible processes through an entropy function, and so on.
+While the ensemble space for each theory will in principle be different, they will all share common features: they must allow measurable properties of the system, probability distributions for outcomes of said properties, characterize reversible and irreversible processes through an entropy function, and so on.
 We are looking for necessary minimal basic requirements upon which these common tools can be built.
 
-At least two arguments can be made to show that any physical theory must allow statistical descriptions, provide an ensemble space.
+At least two arguments can be made to show that any physical theory must allow statistical descriptions, i.e.~provide an ensemble space.
 First, note that physical laws are never about specific instances but rather regularities.
@@
 Second, if a physical theory is to be testable experimentally, and repeatedly so, it must be in terms of ensembles because this is what we prepare in practice and this is what, in the end, we characterize experimentally.
-The approach, then, can most likely be given firm conceptual grounds, which we leave as open but solvable philosophical problem.
+The approach, then, can most likely be given firm conceptual grounds, which we leave as an open but solvable philosophical problem.
 
@@
 \subsection{Experimental verifiability}
-The first requirement is that ensemble must be connected to experimental verification.
+The first requirement is that ensembles must be connected to experimental verification.
 That is, we must have enough \textbf{verifiable statements} at our disposal to define ensembles and tell them apart.
@@
 For example, the statement ``the mass of the photon is less than $10^{-13} \, eV$'' is verifiable, while ``the mass of the photon is exactly $0 \, eV$'' is not verifiable due to its infinite precision.
-Based on previous work, this mathematically requires an ensemble space to be a $T_0$ second countable topological space, where open sets represent verifiable statements and Borel sets represent statements associated with a test, but without guarantee of termination.
+Based on previous work, this mathematically requires an ensemble space to be a $T_0$ second-countable topological space, where open sets represent verifiable statements and Borel sets represent statements associated with a test, but without guarantee of termination.
 
-The topological structure, then, represent the most foundational structure for a physical theory.
+The topological structure, then, represents the most foundational structure for a physical theory.
 It tells us why functions are ``well-behaved,'' which simply means they have to be topologically continuous in the ``natural topology'' induced by this requirement.
@@
 non-Borel sets are physically ill-defined as they are not connected to experimental verification).
-Similarly, it tells us that sets with cardinality greater than that of the continuum are physically not relevant, as they cannot be given a $T_0$ second countable topology.
+Similarly, it tells us that sets with cardinality greater than that of the continuum are not physically relevant, as they cannot be given a $T_0$ second-countable topology.
 
-In the case of ensembles, typical verifiable statements will be ``the average energy of the particle is between 3 and 4 $eV$,'' ``the probability of getting heads is between 49 and 51 percent,'' ``the probability distribution for the position is a Gaussian of average 0 $m$ and 1 $m$ standard deviation within 1\%.'' Note that these are not results of single shot measurements, and therefore verifiable statements are distinctly different from observables in quantum mechanics.
+In the case of ensembles, typical verifiable statements will be ``the average energy of the particle is between 3 and 4 $eV$,'' ``the probability of getting heads is between 49 and 51 percent,'' ``the probability distribution for the position is a Gaussian of mean 0 $m$ and 1 $m$ standard deviation within 1\%.'' Note that these are not results of single-shot measurements, and therefore verifiable statements are distinctly different from observables in quantum mechanics.
 In the same manner, these are not measurements that extract one bit of information, as there is no requirement of termination in the negative case.
@@
 Mathematically, the ensemble space is endowed with a \textbf{convex structure}.
-Moreover, since the mixing operation must be consistent with experimental verifiability, it will be topologically continuous and the ensemble space will be a topological convex space.
+Moreover, since the mixing operation must be consistent with experimental verifiability, it will be topologically continuous, and the ensemble space will be a topological convex space.
 
@@
 The entropy is ultimately responsible for all geometrical structure in physics.
-In classical mechanics, the geometry is essentially defined by phase space volumes and areas in each DOF (degree of freedom).
-The entropy of uniform distributions is given by the logarithm of the volume, meaning that given the volumes we are able to calculate the entropy and given the entropy we are able to reconstruct the volume.
+In classical mechanics, the geometry is essentially defined by phase space volumes and areas in each degree of freedom (DOF).
+The entropy of uniform distributions is given by the logarithm of the volume, meaning that given the volume we are able to calculate the entropy and given the entropy we are able to reconstruct the volume.
 Similarly, uniform marginals will recover DOF areas.
@@
 On the convex structure we can define \textbf{pure states} as the ones that cannot be decomposed into other ensembles.
-For example, the pure states for a six-sided die are the ensemble perfectly prepared respectively on the six possible outcomes.
+For example, the pure states for a six-sided die are the ensembles perfectly prepared respectively on the six possible outcomes.
 All ensembles are convex combinations $\sum_{i=1}^{6} p_i \ens_i = 1$ of those pure states $\ens_i$.
@@
 For example, in classical mechanics the pure states would be perfect preparations of position and momentum, and ensembles are probability distributions.
-Not only this is problematic mathematically, it is conceptually not sustainable.
+Not only is this problematic mathematically, it is conceptually not sustainable.
 
@@
 The mathematical facts that delta Dirac are not proper functions, their support has an empty interior, have minus infinite entropy, and so on, are simply a consequence of our conceptual problem.
-One may concede that the ensembles to do not exist physically, but may still want them to consider them as the limit of infinitesimal decomposition: classical ensembles can be decomposed into more and more precise ones, and the points of are the limit of this decomposition.
+One may concede that the ensembles do not exist physically, but may still want them to consider them as the limit of infinitesimal decomposition: classical ensembles can be decomposed into more and more precise ones, and the points are the limit of this decomposition.
 Probability distributions are defined on these limits.

### 2024-11-12 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Minor typos

@@
 \begin{tcolorbox}[colback=white, colframe=black]
-	Statistical mixture of ensemble $\Rightarrow$ An ensemble space $\Ens$ is a topological convex space.
+	Statistical mixture of ensembles $\Rightarrow$ An ensemble space $\Ens$ is a topological convex space.
 \end{tcolorbox}
@@
 \subsection{Entropy}
-The last requirement is that each ensemble must have a well defined \textbf{entropy}. That is, there is a scalar function $S : \Ens \to \mathbb{R}$ that satisfies the following
+The last requirement is that each ensemble must have a well defined \textbf{entropy}. That is, there is a scalar function $S : \Ens \to \mathbb{R}$ that satisfies the following:
 \begin{enumerate}

### 2024-11-14 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: more work on fraction capacity

@@
 
-Note that the mixing coefficient $p$ is connected to probability but it is not itself probability.
+Note that the mixing coefficient $p$ is connected to probability but it is not itself probability in the standard sense.
 It represents a parameter for the mixing states, mixing preparations, while probability is typically understood as the chance of something happening, the likelihood of a final event.
 Mathematically, the probability measure is a function from an event, which can be understood as a collection of fully distinct outcomes, to a number from zero to one.
-If we are mixing ensemble that represent fully distinct outcomes (i.e.
-they are orthogonal), then the mixing $p$ is a probability.
-If we are mixing ensemble that do not represent fully distinct outcomes (e.g.
-spin up and spin left), the probability of getting one of the outcomes is not the mixing coefficient $p$.
+If we are mixing ensembles that represent fully distinct outcomes (i.e.
+they are orthogonal), then the mixing $p$ is a probability in the standard sense.
+If we are mixing ensembles that do not represent fully distinct outcomes, that are not orthogonal like spin up and spin left, the probability of getting one of the outcomes is not the mixing coefficient $p$.
 
-This is exactly what happens in quantum mechanics: probability is defined only on a set of orthogonal states.
-If we focus only on preparation, then, we only talk about mixing coefficients across non-orthogonal elements and define a more general tool that, when restricted on orthogonal sets, will recover probability.
+We essentially have two ways in which two events can be compared.
+We have \textbf{disjoint events}, those that represent disjoint sets of different states, like spin up and spin down; we also have \textbf{orthogonal events}, those that represent represent orthogonal sets of states, like spin up and spin down.
+The standard notion of probability applies only when considering orthogonal events.
+Classical probability rests on the assumption that all disjoint events are orthogonal.
+That is, that all states can be perfectly distinguished.
+Quantum mechanics breaks this assumption, and therefore we need a probability theory that can deal with disjoint but non-orthogonal events.
 
-We concentrate on the following question: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subset \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
+Let us concentrate on the following question: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
 How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
-For example, if $\ens$ is a uniform distribution over a fair six-sided die, and $A$ contains four ensemble, each fully on ${1,2}$ and uniform distribution over ${5,6}$ two and four, than $5/6$ of $\ens$ can be constructed from $A$, which is the probability that we d
+For example, let $\Ens$ be the space of all probability distributions for a six-sided die.
+Let $\ens_{123456}$ be the uniform distribution over all outcomes.
+Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
+Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$, where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/6$ of $\ens$ can be constructed from $A$, but no more.
+Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
+Note that, given the uniform distribution, $1/6$ is exactly the probably for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
+This gives the basic insight for our definition.
 
-Given a target ensemble $\ens \in \Ens$ and an arbitrary ensemble $\ens[a] \in \Ens$ we will be able to write $\ens = p \ens[a] + (1-p) \ens[b]$ for some $p \in [0,1]$ and some $\ens[b] \in \Ens$.
-At the very least, we can use $p=0$ and $\ens[b] = \ens$.
-Given that $p$ is bounded, there will be a biggest $p$ possible which we call the \textbf{fraction} of $\ens[a]$ in $\ens$.
-Mathematically
+Given a target ensemble $\ens \in \Ens$ and an arbitrary ensemble $\ens[a] \in \Ens$, we define the \textbf{fraction} of $\ens[a]$ in $\ens$ to be
 \begin{equation}
@@
 \end{equation}
-This quantity tells us how much of ensemble $\ens$ can be constructed, can be characterized, by $\ens[a]$.
+The fraction is always well-defined because we can always write $\ens = 0 \ens[a] + 1 \ens$, therefore it must be zero or greater.
+This quantity tells us how much of ensemble $\ens$ can be constructed from $\ens[a]$.
 
-We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a set of ensembles $A \subset \Ens$., meaning we want to characterize how much of a target ensemble $\ens$ can be constructed as a mixture of ensembles from $A$.
-Given a set of ensembles $A$, its \textbf{hull}, noted $\hull(A)$ is the set of all mixtures that can be constructed from $A$.
-That is, is the set of all convex combinations $\sum_i^n p_i \ens[a]_i$ such that $\{a_i\}_1^n \subseteq A$.
-The \textbf{fraction capacity} is the biggest fraction among all the elements of the hull.
-Mathematically
+We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a Borel set of ensembles $A \subset \Ens$.
+Recalling that the $\hull(A)$ is the set of all possible convex combinations (i.e.
+mixtures), we define the\textbf{fraction capacity} of $A$ for $\ens$ to be
 \begin{equation}
@@
 \end{equation}
+This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$.\footnote{The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
+fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.
+TODO \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} }
 
-If $\Ens$ is a topological space, we can imagine the fraction capacity to be defined on the sigma algebra $\Sigma_{\Ens}$.
-We can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
+One can then that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
 \begin{enumerate}
@@
 	\item monotone - $A \subseteq B \implies \frcap_{\ens}(A) \leq \frcap_{\ens}(B)$
-	\item sub-additive - $\frcap_{\ens}(A \cup B) \leq \frcap_{\ens}(A) + \frcap_{\ens}(B)$.
+	\item sub-additive - $\frcap_{\ens}(A \cup B) \leq \frcap_{\ens}(A) + \frcap_{\ens}(B)$
 	\item continuous from below and above - $\frcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \frcap_{\ens}(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
 \end{enumerate}
+
+The first property readily comes from the domain of the fraction.
+For the second, note that $\hull$ and $\sup$ are both monotone.
+For sub-additivity, note that the fraction of two components can at most sum during mixing.
+For the last, increasing and decreasing sequences strictly add or remove possible mixtures, therefore they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
+The limit of these sequences must agree with the fraction capacity of the limit.
+
+Note that a probability satisfies all the properties of a fraction capacity, but it also satisfy on additional property: additivity over disjoint sets.
+That is, $\mu(A \cup B) = \mu(A) + \mu(B)$ if $A \cup B = \emptyset$.
+This is not true in general, as we can see in a simple example from quantum mechanics.
+Let $\Ens$ be the ensemble space of a qubit, that is the space of density matrices of a two dimensional Hilbert space.
+Geometrically, this is the convex set represented by a three dimensional ball.
+Let $\ens_{\psi} = |\psi\>\<\psi|$ be a pure state and let $\ens_{\phi} = |\phi\>\<\phi|$ be its orthogonal.
+Geometrically, these are two opposite points on the surface of the ball.
+The maximally mixed state can be expressed as $\ens = \frac{1}{2} \ens_0 + \frac{1}{2} \ens_1$.
+This is the center point of the ball and is in fact the midpoint between two opposite points.
+Since this is valid for any point, we have $\frcap_{\ens}(\{\ens_{\psi}\})=\frac{1}{2}$ for any $\ens_{\psi} \in \Ens$ and $\frcap_{\ens}(\{\ens_{\psi}, \ens_{\phi}\})=1$ for any pair of orthogonal pure states.
+If $A$ is a set of two orthogonal pure states, and $B$ is state different from the two, we have $\frcap_{\ens}(A \cup B) = 1 < \frcap_{\ens}(A) + \frcap_{\ens}(B) = \frac{3}{2}$.
+
+If we restrict ourselves to lattice of subsets of $\Ens$ we can recover additivity.
+For example, if we take a family $A_i$ of orthogonal sets (i.e.
+all the elements of one set are orthogonal to all the elements of any other set) and take an element that is a mixture of elements from that family, then the fraction capacity will be additive over the lattice formed by closing the family under union and intersection.
+In quantum mechanics, this would correspond to all probability distributions that are outcomes of the projective measurement that distinguishes between $A_i$.
+Part of the task, then, will be finding a set of necessary and sufficient conditions under which the fraction capacity for a particular ensemble is additive over a particular lattice of subsets
+
+What is important here is that
 

### 2024-11-21 Articles/2023-HilbertQuantum/Submissions/Quantum Studies/HilbertQuantum.tex
Updates after Francisco's comments and further proofreading

@@
 \begin{abstract}
-We show that Hilbert spaces are not suitable to represent quantum states mathematically, in the sense that they require properties that are untenable by physical entities.
-We first prove that the requirements posited by complex inner product spaces are physically justified.
-We then show that completeness in the infinite-dimensional case requires the inclusion of states with infinite expectations, coordinate transformations that take finite expectations to infinite ones and vice-versa, and time evolutions that transform finite expectations to infinite ones in finite time.
+We argue that Hilbert spaces are not suitable to represent quantum states mathematically, in the sense that they require properties that are untenable by physical entities.
+We first demonstrate that the requirements posited by complex inner product spaces are physically justified.
+We then show that completeness in the infinite-dimensional case requires the inclusion of states with infinite expectations, coordinate transformations that take finite expectations to infinite ones and vice versa, and time evolutions that transform finite expectations to infinite ones in finite time.
 This means we should be wary of using Hilbert spaces to represent quantum states as they turn a potential infinity into an actual infinity.
@@
 While we do not claim to know what a physically more appropriate closure should be, we note that Schwartz spaces, among other things, guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are closed under Fourier transform.
+
 \end{abstract}
@@
 This means that a truly general statement in quantum theory must be valid also on these irrelevant unphysical cases, with the consequence that either physical results are left out, or they are obscured in a complicated patchwork of theorems with different domains of validity.
-We therefore suspect that some of the mathematical difficulties encountered when trying to give a more precise mathematical foundation to quantum theory stem from this fundamental mismatch.
+We therefore suspect that some of the mathematical difficulties encountered when trying to give a more precise mathematical foundation to quantum theory stem from this fundamental mismatch, and we argue more generally that mathematical spaces used in physics should already come equipped with the proper structure that excludes physically pathological behavior.
 
@@
 We will see that the problem lies in requiring completeness, which should be understood as an unphysical mathematical requirement.
-We will show that other options exist, offering Schwartz spaces as an example.
+We will show that other potential options exist, offering Schwartz spaces as an example.
 These guarantee the expectation values of all polynomials of position and momentum to be finite, and they are also closed under Fourier transform.
@@
 It does not seem to be the case: a reviewer from a previous journal submission stated that ``the question of whether the limiting states are actually physical is rather pointless;'' conversely, others see no problem in working with quantum field theory path integrals that are not mathematically well-defined.
-We are therefore in the absurd, from our perspective, position of having to defend that, yes, the math we need to use in physics has be mathematically precise and, yes, the math we need to use in physics has to represent---perhaps idealized---physically realizable objects.
-The cost of this failure is all around us: an ever-increasing number of interpretations of quantum mechanics, abstract mathematical work whose connection to physics is unclear, theoretical physics more and more disconnected from experimental physics.
-The reader who, like us, finds this obvious can skip this section.
+We are therefore in the absurd, from our perspective, position of having to defend that, yes, the math we need to use in physics has to be mathematically precise and, yes, the math we need to use in physics has to represent---perhaps idealized---physically realizable objects.
+The cost of not requiring a clear and tight connection between the math and the physics is all around us: an ever-increasing number of interpretations of quantum mechanics, abstract mathematical work whose connection to physics is unclear, theoretical physics more and more disconnected from experimental physics.
+The reader who finds this obvious can skip this section.
 
@@
 When we say, ``States are rays in a Hilbert space,'' we are giving a physical specification.
-We are saying that there are objects in the real world that can be modeled with a particular mathematical structure.
-If Hilbert spaces do not work, it is not simply a math problem: the model is wrong, which is a physics problem.
+We are saying that there are objects in the physical world that can be modeled with a particular mathematical structure.
+If Hilbert spaces do not work to model physical states, it is not simply a math problem: the model is wrong, which is a physics problem.
 This is the question we are examining in this paper: does the math we use in quantum mechanics model the physical world correctly?
@@
 To us, and to every experimental physicist we talked to, ``It doesn't make sense experimentally,'' and this is the best argument possible to show that something is unphysical.
-``It can't be done.'' No argument is better.
-However, more than one previous reviewer found no problem with infinite (or undefined) expectation values, as they adopt an ``ensemble interpretation,'' and tried to convince us that we should not be concerned either.
-How can progress on the mathematical foundations of quantum mechanics be made if an argument that is accepted by the vast majority of physicists is rejected when one tries to discuss it in the literature?
+``It can't be done, not even in principle.''
 
@@
 They clearly are.
-The question at hand is whether, on physical grounds, this is a good idea.
+The question at hand is whether, on physical grounds, this is a judicious idea.
 Should we really treat continuous and discrete spectra the same, or should we consider that they are, on physical grounds, two rather different things?
 Should the fact that the energy spectrum is typically discrete for bound states and continuous for unbound states make us wonder if there is something in the differing topologies of the spectra that is physically important?
-Should we then make the difference more prominent in our thinking, rather than trying to construct mathematical tools that treat them the same?
+Should we then make the difference more prominent in our thinking, rather than trying to construct mathematical tools that treat them the same way?
 Similarly, one can take the algebraic route and start with an algebra of observables.
@@
 What, then, is the suitable physically meaningful state space?
-Moreover, note that in both these frameworks, Hilbert spaces still play a crucial role.
+One of the ``successes'' of the algebraic approach was to construct a Hilbert space from an abstract $C^*$-algebra via the so-called GNS construction.
+Thus in both these frameworks, rigged Hilbert spaces and the $C^*$-algebra approach, Hilbert spaces still play a crucial role.
 If Hilbert spaces are physically problematic, these approaches are still relying on a mathematical structure that is physically problematic.
+Looking to a completely different field, within the hierarchy of safety controls, elimination, substitution, and engineering controls are strongly preferred over administrative controls to manage hazards---putting up a fence at the edge of a cliff is much safer than simply putting a sign saying, ``Beware of cliff.''  By using Hilbert spaces for physical calculations in quantum mechanics, the community is effectively relying on ``administrative controls'' to avoid incorrect results, i.e.~considering if a mathematically allowed calculation makes physical sense and discarding it if not.
+For physics problems in realms close to everyday human experience, our physical intuition is very reliable.
+However, for problems within e.g.~quantum physics, quantum field theory, or cosmology, our physical intuition is much less powerful, and in fact these areas of physics rely much more heavily on the mathematics.
+Our paper is effectively advocating that we try to implement ``engineering controls'' to ensure the soundness of calculations in quantum mechanics---if we can find a suitable mathematical space that does not allow unphysical objects in the first place, then there is no risk of accidentally failing to discard unphysical results.
 
@@
 Let us give an example.
-In a previous work we found that a set of physically distinguishable objects must be a $T_0$ second-countable topological space where each open set represents a verifiable statement, a statement allowing a test that terminates in finite time if and only if the statement is true\cite{aop-book}.
+In a previous work some of us found that a set of physically distinguishable objects must be a $T_0$ second-countable topological space where each open set represents a verifiable statement, a statement allowing a test that terminates in finite time if and only if the statement is true\cite{aop-book}.
 This result tells us why functions in physics must be ``well-behaved'' (i.e.~topologically continuous in the right topology), why sets with cardinality greater than the continuum are unphysical (i.e.~they cannot be given a $T_0$ second-countable topology) and why the Banach-Tarski paradox does not apply in physics (i.e.~non-Borel sets are unphysical as they cannot be associated with an experimental test).
@@
 Are Hilbert spaces capturing, and only capturing, physically tenable assumptions, or not?
+%We note that von Neumann himself placed very high value on conceptual clarity in physical theories.\cite{Redei2006}
 
@@
 However, we get a consistent definition if we restrict regions to be Borel sets, instead of arbitrary subsets.
-Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{aop-book} we show that the Borel sets are exactly those sets that can be associated with experimental procedures.
+Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{aop-book} some of us show that the Borel sets are exactly those sets that can be associated with experimental procedures.
 We have often found that physical requirements naturally map to well-behaved spaces.} we need to look more closely at what the math dragged in.
@@
 For example, changes of basis and changes of variables are unitary transformations.
-This is already a problem: the space of wave-functions for $n$ particles is the Hilbert space $L^2(\mathbb{R}^{3n})$.
+This is already a problem for physics: the space of wave functions for $n$ particles is the Hilbert space $L^2(\mathbb{R}^{3n})$.
 For any value of $n$, these infinite-dimensional spaces are separable, meaning that they all have a countable basis, and so are unitarily equivalent.
@@
 
-The second reason for the importance of unitary transformation is that they represent possible time evolution for an isolated system.
+The second reason for the importance of unitary transformations is that they represent possible time evolution for an isolated system.
 The issue is that unitary transformations in a Hilbert space can map states with finite expectation values to those with infinite ones, leading to obvious problems.
@@
 \end{equation}
-This is clearly continuous in $t$ and, for $t=0$, we have the identity transformation, and therefore, mathematically, there is no reason why this wouldn't be a valid time evolution.
-What this does is stretch back and forth the wavefunction, such that the state keeps oscillating between the two functions above, making the expectation value oscillate from finite to infinite in finite time.
+This is clearly continuous in $t$ and, for $t=0$, we have the identity transformation, and therefore, mathematically, there is no reason why this would not be a valid time evolution.
+What this does is stretch back and forth the wave function, such that the state keeps oscillating between the two functions above, making the expectation value oscillate from finite to infinite in finite time.
 This, again, is physically untenable.
@@
 We are likely missing out on useful results by not considering what is provable within the subset of physical cases.
-We, therefore, wonder whether many of the technical issues we have in quantum theory (e.g.
-the non-existence of a valid measure for path integrals\cite{glimm_quantum_1987}, inequivalent representations for interacting theories\cite{earman_haags_2006}, the non-existence of the categorical tensor product for Hilbert spaces\cite{garrett_tensor_2010, sorkin_inside_2022}, etc.)  are simply a consequence of using unphysical objects at a fundamental level.
+We, therefore, wonder whether many of the technical issues we have in quantum theory (e.g.~the non-existence of a valid measure for path integrals\cite{glimm_quantum_1987}, inequivalent representations for interacting theories\cite{earman_haags_2006}, the non-existence of the categorical tensor product for Hilbert spaces\cite{garrett_tensor_2010, sorkin_inside_2022}, etc.)  are simply a consequence of using unphysical objects at a fundamental level.
 
@@
 
-While we do need infinite-dimensional spaces to handle unbound quantities, such as the number of particles, or continuous quantities, such as position, it seems that Cauchy-completeness is not the proper characterization.
+While we do need infinite-dimensional spaces to handle unbound quantities, such as the number of particles, or continuous quantities, such as position, it seems that completeness is not the proper characterization.
 While finding the proper one goes beyond the scope of this work, we want to show that fruitful lines of inquiry do exist.
@@
 In quantum mechanics, this means requiring all expectations of $\frac{1}{2}(X^nP^m + P^mX^n)$ to be finite.
-This restricts our space to those functions that are infinitely smooth (i.e.
-all derivatives exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.
+This restricts our space to those functions that are infinitely smooth (i.e.~all derivatives exist) and decrease very rapidly as $x$ goes to infinity: these are the Schwartz functions.
 
@@
 This appendix presents the same arguments in the article's main body, but expanded with full mathematical detail.
-The figure and the derivations in the Appendix are courtesy of the Assumptions of Physics project.
-\cite{aop-book}
+The figure and the derivations in the Appendix are courtesy of the Assumptions of Physics project.\cite{aop-book}
 
@@
 \end{equation}
-The angle is divided by two because the angle on the Bloch sphere (i.e.
-in physical space) is double the angle in the Hilbert space.
-For example, the angle between $+$ and $-$ on the Bloch sphere is $\pi$, but the inner product between them is zero (i.e.
-opposite directions in physical space correspond to orthogonal states) and therefore the angle in the Hilbert space is $\pi/2$.
+The angle is divided by two because the angle on the Bloch sphere (i.e.~in physical space) is double the angle in the Hilbert space.
+For example, the angle between $+$ and $-$ on the Bloch sphere is $\pi$, but the inner product between them is zero (i.e.~opposite directions in physical space correspond to orthogonal states) and therefore the angle in the Hilbert space is $\pi/2$.
 
@@
 However, because the expectation diverges, the limit cannot be in the domain of $X$.
-Cauchy completeness forces us to include ``states'' for which the position is ill-defined.
+Completeness forces us to include ``states'' for which the position is ill-defined.
 
@@
 More concretely, a change of variable $x$ to $y$ induces a unitary operator as it is just a change of representation.
-The transformation rules for wavefunctions simply follow by the fact that $\psi^\dagger(x) \psi(x) = \rho(x)$ gives us a probability density over $x$.
+The transformation rules for wave functions simply follow by the fact that $\psi^\dagger(x) \psi(x) = \rho(x)$ gives us a probability density over $x$.
 Therefore we must have
@@
 Let $y=y(x)$ be our change of variable such that $y(0) = 0$.
-If $\psi(x)$ is a wavefunction expressed over $x$, and $\phi(y)$ represents the same state as a wavefunction over $y$, we have the following relationship:
+If $\psi(x)$ is a wave function expressed over $x$, and $\phi(y)$ represents the same state as a wave function over $y$, we have the following relationship:
 \begin{equation}
@@
 \end{equation}
-To find a coordinate transformation that takes finite expectations to infinite expectations, we set the initial and final wavefunctions as follows:
+To find a coordinate transformation that takes finite expectations to infinite expectations, we set the initial and final wave functions as follows:
 \begin{align}
@@
 \end{equation}
-The idea is that the evolution will simply stretch and shrink the wavefunction intermittently, potentially oscillating expectation values from finite to infinite and vice versa in finite time.
+The idea is that the evolution will simply stretch and shrink the wave function intermittently, potentially oscillating expectation values from finite to infinite and vice versa in finite time.
 Clearly, this is not a physically meaningful time evolution, yet it will correspond to a strongly continuous one-parameter unitary group and, by Stone's theorem, will even admit a Hamiltonian.
@@
 
-Some physical systems are described by unbounded discrete quantities (e.g.
-number of particles) or continuous quantities (e.g.
-position), which require infinite-dimensional spaces.
+Some physical systems are described by unbounded discrete quantities (e.g.~number of particles) or continuous quantities (e.g.~position), which require infinite-dimensional spaces.
 Infinite-dimensional complex inner product spaces, then, are needed to represent quantum systems.
@@
 At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true for distributions in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
-Therefore the space of distributions with compact support on phase space gives us finite expectation value for all functions of positions and momentum.}
+Therefore the space of distributions with compact support on phase space gives us finite expectation value for all functions of position and momentum.}
 
@@
 In quantum mechanics, this means requiring all expectations of $X^nP^m = -\imath \hbar x^n\partial_x ^m$ to be finite.
-This restricts our space to those functions that are infinitely smooth (i.e.
-all derivatives exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions.
+This restricts our space to those functions that are infinitely smooth (i.e.~all derivatives exist) and decrease very rapidly as $x$ goes to infinity: the Schwartz functions.
 Requiring finite expectation for all polynomials of position and momentum, then, does not give us the usual $L^2$ Hilbert space; it gives us the Schwartz space.
@@
 The Schwartz space is, in fact, a Fréchet space \cite[Theorem V.9]{reed_methods_1980}, meaning that it is a Hausdorff space, with the topology induced by the seminorms given by the expectations, and it is complete with respect to those seminorms.
-Physically, this means that measurements on the polynomials (i.e.
-the topology of the seminorms) are enough to identify the states (i.e.
-the singletons, which are closed sets as per Hausdorff).\footnote{This essentially tells us that the moment problem restricted to pure states has a unique solution in quantum mechanics.
+Physically, this means that measurements on the polynomials (i.e.~the topology of the seminorms) are enough to identify the states (i.e.~the singletons, which are closed sets as per Hausdorff).\footnote{This essentially tells us that the moment problem restricted to pure states has a unique solution in quantum mechanics.
 Setting up a comparison with the classical case is delicate.
-If we restrict ourselves to classical pure states, i.e.
-points in phase space, then the solution is also unique.
+If we restrict ourselves to classical pure states, i.e.~points in phase space, then the solution is also unique.
 If we restrict ourselves to compact support, distinct distributions have distinct moments even in the multi-variable case.\cite{moment_problem_2017} If we restrict ourselves to integrable functions, even in the single variable case, distinct distributions can have the same moments.
 None of these comparisons seems adequate as a quantum state is a statistical object by nature, but it is not an arbitrary mixture.
-A more proper comparison would seem to restrict the problem to the space of integrable functions with zero entropy.} Moreover, if we have a sequence of states for which all expectations converge, we are guaranteed to converge to another state (i.e.
-completeness with respect to the seminorms).
+A more proper comparison would seem to restrict the problem to the space of integrable functions with zero entropy.} Moreover, if we have a sequence of states for which all expectations converge, we are guaranteed to converge to another state (i.e.~completeness with respect to the seminorms).
 If any expectation diverges, the limit is not in the space.
@@
 That is, the Fourier transform of a Schwartz function is a Schwartz function.
-Physically, this means that we can always express a wavefunction in either the position or the momentum basis.
+Physically, this means that we can always express a wave function in either the position or the momentum basis.
 Mathematically, the theory of tempered distributions, which allows us to define objects like the Dirac delta distribution, is built on top of the Schwartz space.
@@
 
-Once mathematicians and philosophers noticed that mathematical theories allowed one to express inconsistencies (e.g.
-Russell's paradox, the Berry paradox, etc.), an effort was put into strengthening the foundations of mathematics so that such paradoxes were either resolved or couldn't be expressed.
+Once mathematicians and philosophers noticed that mathematical theories allowed one to express inconsistencies (e.g.~Russell's paradox, the Berry paradox, etc.), an effort was put into strengthening the foundations of mathematics so that such paradoxes were either resolved or couldn't be expressed.
 Concerning the mathematical foundations of physical theories, we have a similar problem: we can write statements that, though mathematically consistent, are physically inconsistent.

### 2024-11-26 Articles/2023-HilbertQuantum/Submissions/Quantum Studies/HilbertQuantum.tex
Remove awkward sentence about von Neumann
Final revision after discussion with Francisco; resubmission version

@@
 But mathematics plays a much more fundamental role in physics: it is the language in which we specify our models.
-When we say, ``States are rays in a Hilbert space,'' we are giving a physical specification.
-We are saying that there are objects in the physical world that can be modeled with a particular mathematical structure.
-If Hilbert spaces do not work to model physical states, it is not simply a math problem: the model is wrong, which is a physics problem.
-This is the question we are examining in this paper: does the math we use in quantum mechanics model the physical world correctly?
+When we say, ``States are rays in a Hilbert space,'' we are providing a specification for a physical system.
+We are saying that there is a particular physical system which under certain circumstances can be mathematically modeled in a particular way.
+So, if Hilbert spaces do not work, it is not simply a math problem: the model is wrong, which is a physics problem.
+This is the question we are examining in this paper: can the math we use in quantum mechanics model a particular class of physical systems correctly?
 Is our physical problem well specified?
@@
 Are Hilbert spaces capturing, and only capturing, physically tenable assumptions, or not?
-%We note that von Neumann himself placed very high value on conceptual clarity in physical theories.\cite{Redei2006}
 

### 2024-11-26 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
ClassicalLimit: addressing comments

@@
 
+We focus on maps that increase the entropy uniformly because, in the end, we will want Hamiltonian evolution on the original states to correspond to Hamiltonian evolution on the transformed states.
+Since Hamiltonian evolution maps states at constant entropy, $S(R(\rho))$ must be a function of $S(\rho)$.
+Moreover, note that for any $\rho_1$ and $\rho_2$ with disjoint support, $S\left(\frac{1}{2}\rho_1 + \frac{1}{2}\rho_2\right) = 1 + \frac{1}{2}S\left(\rho_1\right) + \frac{1}{2}S\left(\rho_2\right)$.
+Since $R$ will preserve disjointness of support, the one bit of entropy increase will have to carry through the map, meaning that the difference in entropy will need to be preserved.
+This is why we are focusing on maps that increase the entropy uniformly.
+
 For a generic transformation, the increase of entropy is given by the expectation of the Jacobian determinant $\Delta S =\int_M \rho \log |\partial_a R^b| dxdp$.
@@
 
+Note that while entropy could be increased in many different ways, under a stretching map unitary evolution will be mapped to another unitary evolution.
+In fact, if $H$ is a polynomial of position and momentum, we will be able to write the corresponding $\hat{H}$ such that $T_Q(\hat{H}) = H$.
+The stretching map, then, is useful to show not only that high-entropy states can be approximated by classical states, but that unitary evolution can be approximated by classical Hamiltonian mechanics.
+One may show that arbitrary entropy-increasing maps will recover classical states, for example by applying similar techniques to entropy bounds\cite{Hall_2018}, but they will not necessarily recover classical Hamiltonian evolution.
+
 \begin{figure}[h]

### 2024-11-27 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
Proofread before submission to a journal

@@
 
-The goal of this paper, summarized in figure \ref{fig_quadrants}, is to show that classical mechanics is the high-entropy limit of quantum mechanics, much in the same way that it is the low-speed limit of relativistic mechanics.
+The goal of this paper, summarized in Fig.~\ref{fig_quadrants}, is to show that classical mechanics is the high-entropy limit of quantum mechanics, much in the same way that it is the low-speed limit of relativistic mechanics.
 
@@
 This way, all finite entropies are large compared to the lowest possible entropy.
-In a nutshell, this is what we will show in section four.
+In a nutshell, this is what we will show in section 4.
 Since physically it does not make sense to take limits of physical constants, $c \to \infty$ should be understood as $v \ll c$, $\hbar \to 0$ should be understood as $S \gg 0$.
@@
 It has been shown that coherence can be maintained over long distances and among a large number of constituents, meaning that quantum systems are not necessarily small or made of few components.
-However, it is also established how coherence can be quickly lost through interaction with the environment, through decoherence.\cite{RevModPhys.75.715, SCHLOSSHAUER20191} Since decoherence increases the entropy of the system, it represents one mechanism to reach the high-entropy limit.
+However, it is also established how coherence can be quickly lost through interaction with the environment, through decoherence~\cite{RevModPhys.75.715, SCHLOSSHAUER20191}.
+Since decoherence increases the entropy of the system, it represents one mechanism to reach the high-entropy limit.
 This is in line with our result.
@@
 Some quantum effects can be replicated at higher temperature given a high pressure.
-This is the case, for example, in some superconductive materials.\cite{Drozdov:2015} High pressure corresponds to low entropy, which is in line with our result.
+This is the case, for example, in some superconductive materials~\cite{Drozdov:2015}.
+High pressure corresponds to low entropy, which is in line with our result.
 Along the same lines, white dwarfs and neutron stars exhibit quantum effects at high pressure, despite high temperatures.
@@
 
-For a trapped condensate, the PSD at the center and entropy $S$ are related by $\text{PSD}=e^{5/2+\gamma-S/N}$, where $N$ is the number of particles and $\gamma$ is the virial coefficient for the trapping potential.\cite{PhysRevLett.78.990} This has been used experimentally to condense at constant entropy, by changing the shape of the potential.\cite{PhysRevLett.81.2194} Note that, however, the system is inhomogeneous, so the condensation happens because entropy is transferred from one part of the system to the other.
+For a trapped condensate, the PSD at the center and entropy $S$ are related by $\text{PSD}=e^{5/2+\gamma-S/N}$, where $N$ is the number of particles and $\gamma$ is the virial coefficient for the trapping potential~\cite{PhysRevLett.78.990}.
+This has been used experimentally to condense at constant entropy, by changing the shape of the potential~\cite{PhysRevLett.81.2194}.
+Note that, however, the system is inhomogeneous, so the condensation happens because entropy is transferred from one part of the system to the other.
 
@@
 In both classical and quantum mechanics, Gaussian states maximize entropy for a single independent degree of freedom at a fixed uncertainty or, equivalently, minimize uncertainty at fixed entropy.
-This means that, if we fix the entropy $S$, we will have an uncertainty relationship\cite{Carcassi:2022bpm}
+This means that, if we fix the entropy $S$, we will have an uncertainty relationship~\cite{Carcassi:2022bpm}
 \begin{equation}
@@
 
-In classical mechanics, the relationship between entropy and uncertainty for Gaussian states is\cite{Cover_Thomas_2006,Pathria_Beale_2022}
+In classical mechanics, the relationship between entropy and uncertainty for Gaussian states is~\cite{Cover_Thomas_2006,Pathria_Beale_2022}
 \begin{equation}
@@
 \end{equation}
-In quantum mechanics it is\cite{weedbrook2012gaussian}
+In quantum mechanics it is~\cite{weedbrook2012gaussian}
 \begin{equation}
@@
 
-In figure \ref{fig:uncertainty} we can see that the classical and quantum cases are very close even when the uncertainty is just a few units of $\hbar$.
+In Fig.~\ref{fig:uncertainty} we can see that the classical and quantum cases are very close even when the uncertainty is just a few units of $\hbar$.
 Things diverge at about two units of $\hbar$: in quantum mechanics the entropy decreases faster and reaches zero at $\hbar/2$, the bound for the Heisenberg uncertainty principle; classical mechanics reaches zero entropy for $\hbar/e$, and then continues in the negative region.
@@
 Mathematically, the trace distance $T(\rho,\sigma)$ between two mixed states $\rho$ and $\sigma$ is contractive under a completely positive trace-preserving (CPTP) map $M$.
-That is, $T(M(\rho),M(\sigma)) \leq T(\rho,\sigma)$.\cite{nielsen2010quantum} In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive.
-In section 4, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite dimensional case, one could have infinite dimensional subspaces that remain orthogonal while increasing the entropy of all its elements.
+That is, $T(M(\rho),M(\sigma)) \leq T(\rho,\sigma)$~\cite{nielsen2010quantum}.
+In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state~\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive.
+In section 4, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite-dimensional case, one could have infinite-dimensional subspaces that remain orthogonal while increasing the entropy of all its elements.
 This corresponds, for example, to the case where there are conserved quantities.
@@
 }  If we call $\mathcal{G}$ the space of all mixtures of Gaussian states, the states with a non-negative Wigner function, given an arbitrary mixed state $\rho$, we will find a state $\sigma_{\rho} \in \mathcal{G}$ that is closest to $\rho$.
-As we apply our entropy-increasing map, the entropy of $\rho$ increases and $\sigma_{\rho}$ will be get closer to $\rho$.
+As we apply our entropy-increasing map, the entropy of $\rho$ increases and $\sigma_{\rho}$ will get closer to $\rho$.
 In this sense the states will ``look classical.''
@@
 We can similarly reinterpret results that take the mathematical limit $\hbar \to 0$.
-For example, following the original paper from Wigner \cite{WignerLimit}, we can consider the Wigner distribution of a system in thermal equilibrium at inverse temperature $\beta$:
+For example, following the original paper from Wigner~\cite{WignerLimit}, we can consider the Wigner distribution of a system in thermal equilibrium at inverse temperature $\beta$:
 \begin{equation}
@@
 \begin{equation}
-    W(x,p)=\int dy \langle x+y |e^{-\beta\tilde{H}}|x-y \rangle 
+    W(x,p)=\int dy \langle x+y |e^{-\beta\tilde{H}}|x-y \rangle. 
 \end{equation}
@@
 Having given qualitative reasons for the limit, and having seen how, in specific cases, standard arguments can be reinterpreted consistently with our view, we now develop the limit as a direct consequence of entropy increase.
-Since the maximum entropy attainable by a quantum system is the logarithm of the dimension of the corresponding Hilbert space,\cite{nielsen2010quantum} the high-entropy limit exists only for infinite-dimensional spaces.
+Since the maximum entropy attainable by a quantum system is the logarithm of the dimension of the corresponding Hilbert space~\cite{nielsen2010quantum}, the high-entropy limit exists only for infinite-dimensional spaces.
 This explains why spin, which lives in a finite-dimensional space, is an intrinsically quantum property.
@@
 The first takes the product of $n$ operators by averaging all possible permutations; the second is in terms of polynomials of the form $(a^\dagger)^n a^m$, where $a=\sqrt{\frac{m\omega}{2\hbar}}(X+\frac{i}{m\omega}P)$ and $a^\dagger$ are the ladder operators, $m$ is the mass and $\omega$ the angular frequency; the last is in terms of $a^n (a^\dagger)^m$.
-The three orderings are associated, respectively, with a quasi-probability distribution:\cite{carmichael2013statistical} the Wigner function $W$, the Glauber-Sudarshan $P$ distribution and the Husimi $Q$ distribution.
+The three orderings are associated, respectively, with a quasi-probability distribution~\cite{carmichael2013statistical}: the Wigner function $W$, the Glauber-Sudarshan $P$ distribution and the Husimi $Q$ distribution.
 In each case, the expectation under the quasi-probability distribution of polynomials of classical variables returns the expectation of the respective ordering.
@@
 \begin{equation}
-\langle T_W(\Pi(\underbrace{X, ..., X}_{\text{n times}}, \underbrace{P, ..., P}_{\text{m times}})) \rangle = (\sqrt{\lambda})^{(n+m)} \langle \Pi(\underbrace{X, ..., X}_{\text{n times}}, \underbrace{P, ..., P}_{\text{m times}})\rangle.
+\langle T_W(\Pi(\underbrace{X, ..., X}_{\text{n times}}, \underbrace{P, ..., P}_{\text{m times}})) \rangle = (\sqrt{\lambda})^{(n+m)} \langle \Pi(\underbrace{X, ..., X}_{\text{n times}}, \underbrace{P, ..., P}_{\text{m times}})\rangle,
 \end{equation}
@@
 However, this cannot work.
-Wigner functions can have regions with negative values, but the size of these regions cannot exceed a few units of $\hbar$.\cite{kenfack2004negativity} The size of these negative regions would increase under $T_W$, giving functions that do not correspond to a quantum state.
+Wigner functions can have regions with negative values, but the size of these regions cannot exceed a few units of $\hbar$~\cite{kenfack2004negativity}.
+The size of these negative regions would increase under $T_W$, giving functions that do not correspond to a quantum state.
 Therefore we cannot find a pure stretching map in the symmetrized average operator ordering.
@@
 
-We turn our attention to the anti-normal ordering and a map $T_Q$ such that:
+We turn our attention to the anti-normal ordering and a map $T_Q$ such that
 \begin{equation}
@@
 \begin{equation}
-	\frac{d}{d t} a^n (a^{\dagger})^m=\frac{\gamma}{2}(n+m)a^n (a^{\dagger})^m
+	\frac{d}{d t} a^n (a^{\dagger})^m=\frac{\gamma}{2}(n+m)a^n (a^{\dagger})^m,
 \end{equation}
@@
 
-Lastly, we need to show that $T_Q$ increases entropy for all states. Note that a Lindblad operator increases entropy for all states if\cite{Benatti1988}
+Lastly, we need to show that $T_Q$ increases entropy for all states. Note that a Lindblad operator increases entropy for all states if~\cite{Benatti1988}
 \begin{equation}
@@
 As we take the limit $\lambda \to \infty$ in the first viewpoint, in the second viewpoint the commutator $[\hat{X},\hat{P}] = \frac{\imath \hbar}{\lambda}$ decreases, which is equivalent to taking the limit $\hbar \to 0$.
-This is the group contraction that morphs the Moyal bracket Lie algebra to the Poisson bracket Lie algebra.\cite{Moyal_1949,saletan1961contraction, inonu1953contraction} That is, as $\lambda$ increases, the entropy of pure states becomes lower and lower, and the structure of the quantum space becomes closer and closer to the structure of classical mechanics.
+This is the group contraction that morphs the Moyal bracket Lie algebra to the Poisson bracket Lie algebra~\cite{Moyal_1949,saletan1961contraction, inonu1953contraction}.
+That is, as $\lambda$ increases, the entropy of pure states becomes lower and lower, and the structure of the quantum space becomes closer and closer to the structure of classical mechanics.
 
@@
 The stretching map, then, is useful to show not only that high-entropy states can be approximated by classical states, but that unitary evolution can be approximated by classical Hamiltonian mechanics.
-One may show that arbitrary entropy-increasing maps will recover classical states, for example by applying similar techniques to entropy bounds\cite{Hall_2018}, but they will not necessarily recover classical Hamiltonian evolution.
+One may show that arbitrary entropy-increasing maps will recover classical states, for example by applying similar techniques to entropy bounds~\cite{Hall_2018}, but they will not necessarily recover classical Hamiltonian evolution.
 
@@
 The entropy will become $S_{\lambda Q}(\Sigma) = S_Q(\lambda \Sigma) - \ln(\lambda)$.
-As we can see in figure \ref{fig:uncertainty_scaled}, even with low factors of $\lambda$ the entropy curve approaches the classical one.
+As we can see in Fig.~\ref{fig:uncertainty_scaled}, even with low factors of $\lambda$ the entropy curve approaches the classical one.
 If one takes the limit $\lambda \to \infty$, the above expression reduces to the classical equation.
@@
 To do that, we can use the fact that $Q$ is the Weierstrass transform of $W$.
-That is:
+That is,
 \begin{equation}
@@
 \end{equation}
-We notice that these two equations are scaled convolutions between the Wigner distribution and Gaussian functions. Symbolically:
+We notice that these two equations are scaled convolutions between the Wigner distribution and Gaussian functions. Symbolically,
 \begin{equation}
-    (W_\lambda \ast G)(\alpha)=\frac{1}{\lambda}(W_1\ast G)\left(\frac{\alpha}{\sqrt{\lambda}}\right)
+    (W_\lambda \ast G)(\alpha)=\frac{1}{\lambda}(W_1\ast G)\left(\frac{\alpha}{\sqrt{\lambda}}\right),
 \end{equation}
@@
 \begin{equation}
-    W_\lambda(\beta)=F^{-1}\left(F\left(\frac{1}{\lambda}W_1\left(\frac{\beta}{\sqrt{\lambda}}\right)\right)F\left(\frac{1}{\lambda-1}G\left(\frac{\beta}{\sqrt{\lambda-1}}\right)\right)\right)=\frac{1}{\lambda(\lambda-1)}W_1\left(\frac{\beta}{\sqrt{\lambda}}\right) \ast G\left(\frac{\beta}{\sqrt{\lambda-1}}\right)
+    W_\lambda(\beta)=F^{-1}\left(F\left(\frac{1}{\lambda}W_1\left(\frac{\beta}{\sqrt{\lambda}}\right)\right)F\left(\frac{1}{\lambda-1}G\left(\frac{\beta}{\sqrt{\lambda-1}}\right)\right)\right)=\frac{1}{\lambda(\lambda-1)}W_1\left(\frac{\beta}{\sqrt{\lambda}}\right) \ast G\left(\frac{\beta}{\sqrt{\lambda-1}}\right).
     \label{Fourier}
 \end{equation}
-Writing this down explicitly
+Writing this down explicitly,
 \begin{equation}
@@
     \frac{d}{dt} W &= \{\{H, W\}\} \\
-    &= \{H, W\} + \sum_n \frac{\hbar^{2n} (-1)^n}{(2n+1)! \, 2^{2n}} \partial_x^{(2n+1)} V \partial_p^{(2n+1)} W
+    &= \{H, W\} + \sum_n \frac{\hbar^{2n} (-1)^n}{(2n+1)! \, 2^{2n}} \partial_x^{(2n+1)} V \partial_p^{(2n+1)} W.
 \end{aligned}
 \end{equation}
-The evolution is in terms of the Moyal bracket $\{\{H,W\}\}$ which can be expanded in orders of $\hbar$ with the Poisson bracket $\{H,W\}$ as the leading term.
+The evolution is in terms of the Moyal bracket $\{\{H,W\}\}$, which can be expanded in orders of $\hbar$ with the Poisson bracket $\{H,W\}$ as the leading term.
 In our limit, $\hbar$ is constant but as the function stretches all derivatives of $W$ decrease.

### 2024-12-03 Articles/2024-ClassicalLimit/Submissions/SciPost Physics/classicalLimitArticleSciPostPhys.tex
Updated arxiv version and submission version to SciPost Physics

--- 2024-11-27 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
@@
-\documentclass{article}
-\usepackage{authblk}
+% =========================================================================
+% SciPost LaTeX template
+% Version 2024-07
+%
+% Submissions to SciPost Journals should make use of this template.
+%
+% INSTRUCTIONS: simply look for the `TODO:' tokens and adapt your file.
+% ========================================================================
 
-% Language setting
-% Replace `english' with e.g. `spanish' to change the document language
-\usepackage[english]{babel}
-\usepackage{ulem}
-% Set page size and margins
-% Replace `letterpaper' with `a4paper' for UK/EU standard size
-\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
+\documentclass{SciPost}
 
-% Useful packages
-\usepackage{assumptionsofphysics}
-\usepackage[dvipsnames]{xcolor}
+% Prevent all line breaks in inline equations.
+\binoppenalty=10000
+\relpenalty=10000
+
+\hypersetup{
+    colorlinks,
+    linkcolor={red!50!black},
+    citecolor={blue!50!black},
+    urlcolor={blue!80!black}
+}
+
+\usepackage[bitstream-charter]{mathdesign}
+\urlstyle{same}
+
+%\usepackage[dvipsnames]{xcolor}  
+%Had package options conflict.  Added option [dvipsnames] to SciPost.cls file
 \usepackage{tikz-cd}
-\usepackage[colorlinks=true, allcolors=blue]{hyperref}
 \usepackage{pgfplots}
 
-\def\>{\rangle}
-\def\<{\langle}
+% Fix \cal and \mathcal characters look (so it's not the same as \mathscr)
+\DeclareSymbolFont{usualmathcal}{OMS}{cmsy}{m}{n}
+\DeclareSymbolFontAlphabet{\mathcal}{usualmathcal}
 
-\title{Classical mechanics as the high-entropy limit \\ of quantum mechanics}
-\author[1]{Gabriele Carcassi}
-\author[2]{Manuele Landini}
-\author[1]{Christine A. Aidala}
-
-\affil[1]{Physics Department, University of Michigan, Ann Arbor, Michigan 48109, USA}
-
-\affil[2]{Institut f\"ur Experimentalphysik und Zentrum f\"ur Quantenphysik, \protect\\ Universit\"at Innsbruck, Innsbruck, Austria}
+\fancypagestyle{SPstyle}{
+\fancyhf{}
+\lhead{\colorbox{scipostblue}{\bf \color{white} ~SciPost Physics }}
+\rhead{{\bf \color{scipostdeepblue} ~Submission }}
+\renewcommand{\headrulewidth}{1pt}
+\fancyfoot[C]{\textbf{\thepage}}
+}
 
 \begin{document}
-\maketitle
 
-\begin{abstract}
+\pagestyle{SPstyle}
+
+\begin{center}{\Large \textbf{\color{scipostdeepblue}{
+Classical mechanics as the high-entropy limit \\ of quantum mechanics}}}\end{center}
+
+\begin{center}\textbf{
+%%%%%%%%%% TODO: AUTHORS
+% Write the author list here. 
+% Use (full) first name (+ middle name initials) + surname format.
+% Separate subsequent authors by a comma, omit comma and use "and" for the last author.
+% Mark the corresponding author(s) with a superscript symbol in this order
+% \star, \dagger, \ddagger, \circ, \S, \P, \parallel, ...
+Gabriele Carcassi\textsuperscript{1},
+Manuele Landini\textsuperscript{2} and
+Christine A. Aidala\textsuperscript{1$\star$}
+%%%%%%%%%% END TODO: AUTHORS
+}\end{center}
+
+\begin{center}
+%%%%%%%%%% TODO: AFFILIATIONS
+% Write all affiliations here.
+% Format: institute, city, country
+{\bf 1} Physics Department, University of Michigan, Ann Arbor, Michigan 48109, USA
+\\
+{\bf 2} Institut f\"ur Experimentalphysik und Zentrum f\"ur Quantenphysik, \protect\\ Universit\"at Innsbruck, Innsbruck, Austria
+%%%%%%%%%% END TODO: AFFILIATIONS
+%%%%%%%%%% TODO: EMAIL
+% Provide email address of corresponding author(s)
+\\[\baselineskip]
+$\star$ \href{mailto:caidala@umich.edu}{\small caidala@umich.edu}
+
+%%%%%%%%%% END TODO: EMAIL
+\end{center}
+
+\section*{\color{scipostdeepblue}{Abstract}}
+\textbf{\boldmath{%
 We show that classical mechanics can be recovered as the high-entropy limit of quantum mechanics.
@@
 Consequently, the limit presented is independent of possible underlying mechanisms and of what interpretation is chosen for both quantum states and entropy.
-\end{abstract}
+}}
+
+\vspace{\baselineskip}
+
+%%%%%%%%%% BLOCK: Copyright information
+% This block will be filled during the proof stage, and finilized just before publication.
+% It exists here only as a placeholder, and should not be modified by authors.
+\noindent\textcolor{white!90!black}{%
+\fbox{\parbox{0.975\linewidth}{%
+\textcolor{white!40!black}{\begin{tabular}{lr}%
+  \begin{minipage}{0.6\textwidth}%
+    {\small Copyright attribution to authors. \newline
+    This work is a submission to SciPost Physics. \newline
+    License information to appear upon publication. \newline
+    Publication information to appear upon publication.}
+  \end{minipage} & \begin{minipage}{0.4\textwidth}
+    {\small Received Date \newline Accepted Date \newline Published Date}%
+  \end{minipage}
+\end{tabular}}
+}}
+}
+%%%%%%%%%% BLOCK: Copyright information
+
+%%%%%%%%%% TODO: LINENO
+% For convenience during refereeing we turn on line numbers:
+\linenumbers
+% You should run LaTeX twice in order for the line numbers to appear.
+%%%%%%%%%% END TODO: LINENO
+
+%%%%%%%%%% TODO: TOC 
+% Guideline: if your paper is longer that 6 pages, include a TOC
+% To remove the TOC, simply cut the following block
+\vspace{10pt}
+\noindent\rule{\textwidth}{1pt}
+\tableofcontents
+\noindent\rule{\textwidth}{1pt}
+\vspace{10pt}
+%%%%%%%%%% END TODO: TOC
+
+%%%%%%%%% TODO: CONTENTS 
+% Write your article contents here, starting from first \section.
+% An example structure is given below.
 
 \section{Introduction}
+\label{sec:intro}
 
@@
 This way, all finite entropies are large compared to the lowest possible entropy.
-In a nutshell, this is what we will show in section 4.
+In a nutshell, this is what we will show in section~\ref{sec:high_entropy_limit}.
 Since physically it does not make sense to take limits of physical constants, $c \to \infty$ should be understood as $v \ll c$, $\hbar \to 0$ should be understood as $S \gg 0$.
@@
 
-For a trapped condensate, the PSD at the center and entropy $S$ are related by $\text{PSD}=e^{5/2+\gamma-S/N}$, where $N$ is the number of particles and $\gamma$ is the virial coefficient for the trapping potential~\cite{PhysRevLett.78.990}.
+For a trapped condensate, the PSD at the center and entropy $S$ are related by $\text{PSD}=$ $e^{5/2+\gamma-S/N}$, where $N$ is the number of particles and $\gamma$ is the virial coefficient for the trapping potential~\cite{PhysRevLett.78.990}.
 This has been used experimentally to condense at constant entropy, by changing the shape of the potential~\cite{PhysRevLett.81.2194}.
@@
 Mathematically, the trace distance $T(\rho,\sigma)$ between two mixed states $\rho$ and $\sigma$ is contractive under a completely positive trace-preserving (CPTP) map $M$.
-That is, $T(M(\rho),M(\sigma)) \leq T(\rho,\sigma)$~\cite{nielsen2010quantum}.
+That is, $T(M(\rho),M(\sigma)) \leq$ $T(\rho,\sigma)$~\cite{nielsen2010quantum}.
 In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state~\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive.
-In section 4, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite-dimensional case, one could have infinite-dimensional subspaces that remain orthogonal while increasing the entropy of all its elements.
+In section~\ref{sec:high_entropy_limit}, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite-dimensional case, one could have infinite-dimensional subspaces that remain orthogonal while increasing the entropy of all its elements.
 This corresponds, for example, to the case where there are conserved quantities.
@@
 \section{The high-entropy limit}
+\label{sec:high_entropy_limit}
 
@@
 
-\bibliographystyle{plain}
-\bibliography{bibliography}
+\section*{Acknowledgements}
+This paper is part of the ongoing Assumptions of Physics project~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+
+% TODO: include funding information
+\paragraph{Funding information}
+ML acknowledges funding via the FWF-funded QuantERA project QuSiED with project number I 6008-N.
+
+%%%%%%%%% END TODO: CONTENTS
+
+% Use your bibtex library, formatted by the SciPost style file.
+\bibliography{bibliography.bib}
+
+%%%%%%%%%% END TODO: BIBLIOGRAPHY
 

### 2024-12-22 Articles/2023-HilbertQuantum/Submissions/Quantum Studies/HilbertQuantum.tex
Fix references and typo, add acknowledgements
After acceptance by Quantum Studies

@@
 The main point of the paper, then, is to raise awareness that results that rely on the completeness of Hilbert spaces may not be physically significant.
-While we do not claim to know what a physically more appropriate closure should be, we note that Schwartz spaces, among other things, guarantee that the expectation of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are closed under Fourier transform.
+While we do not claim to know what a physically more appropriate closure should be, we note that Schwartz spaces, among other things, guarantee that the expectations of all polynomials of position and momentum are finite, their elements are uniquely identified by these expectations, and they are closed under Fourier transform.
 
@@
 
-Hilbert spaces have been the cornerstone of quantum mechanics since von Neumann's 1932 book.\cite{von_neumann_mathematische_1996} While the formalism is mathematically consistent and most accept it without question, it does not have a clear physical justification.\cite{heathcote_1990, hardy_2001} In fact, von Neumann himself expressed his dissatisfaction not long after publication of his book, and spent considerable effort looking for alternatives.\cite{vonNeumannHilbert_1996}
+Hilbert spaces have been the cornerstone of quantum mechanics since von Neumann's 1932 book~\cite{von_neumann_mathematische_1996}.
+While the formalism is mathematically consistent and most accept it without question, it does not have a clear physical justification~\cite{heathcote_1990, hardy_2001}.
+In fact, von Neumann himself expressed his dissatisfaction not long after publication of his book, and spent considerable effort looking for alternatives~\cite{vonNeumannHilbert_1996}.
 
@@
 
-To make this determination, one needs a standard to judge the physicality of a formal definition\cite{redei_tension_2020, north_physics_2021} and, rather tellingly, physics tradition does not provide one.
-Here we describe the standard that we follow in our work.\cite{aop-book}  Suppose we are studying a physical system.
+To make this determination, one needs a standard to judge the physicality of a formal definition~\cite{redei_tension_2020, north_physics_2021} and, rather tellingly, physics tradition does not provide one.
+Here we describe the standard that we follow in our work~\cite{aop-book}.
+Suppose we are studying a physical system.
 A mathematical definition is \textbf{physical} if it properly characterizes the physical system in the given conditions.
@@
 If the problem of justifying the use of mathematical structures in physics is, ultimately, a physical problem, it also follows that it cannot be solved simply by a different mathematical approach.
-For example, rigged Hilbert spaces were developed to give a mathematical foundation to Dirac’s bra-ket formalism\cite{Dirac1947}, and to be able to talk about eigenstates of position and momentum, working mathematically with both continuous and discrete bases on an interchangeable level\cite{Madrid_2005,Celeghini2016}.
+For example, rigged Hilbert spaces were developed to give a mathematical foundation to Dirac’s bra-ket formalism~\cite{Dirac1947}, and to be able to talk about eigenstates of position and momentum, working mathematically with both continuous and discrete bases on an interchangeable level~\cite{Madrid_2005,Celeghini2016}.
 The question is not whether these can be useful at a mathematical level to solve certain problems.
@@
 Let us give an example.
-In a previous work some of us found that a set of physically distinguishable objects must be a $T_0$ second-countable topological space where each open set represents a verifiable statement, a statement allowing a test that terminates in finite time if and only if the statement is true\cite{aop-book}.
+In a previous work some of us found that a set of physically distinguishable objects must be a $T_0$ second-countable topological space where each open set represents a verifiable statement, a statement allowing a test that terminates in finite time if and only if the statement is true~\cite{aop-book}.
 This result tells us why functions in physics must be ``well-behaved'' (i.e.~topologically continuous in the right topology), why sets with cardinality greater than the continuum are unphysical (i.e.~they cannot be given a $T_0$ second-countable topology) and why the Banach-Tarski paradox does not apply in physics (i.e.~non-Borel sets are unphysical as they cannot be associated with an experimental test).
-This tells us that, at the very least, non-separable Hilbert spaces are unphysical as they have cardinality greater than the continuum.\cite{pittphilsci18363} This ties together ideas from point-set topology, logic, set theory, measure theory and computer science.
+This tells us that, at the very least, non-separable Hilbert spaces are unphysical as they have cardinality greater than the continuum~\cite{Earman2020}.
+This ties together ideas from point-set topology, logic, set theory, measure theory and computer science.
 Clearly, one needs to have at least a basic understanding of these fields to be able to follow the argument, and most degree programs in physics do not typically discuss these topics in depth.\footnote{A previous reviewer conflated the completeness of the Hilbert space with the existence of a complete orthonormal basis of said space.} However, in the end, only a physicist can judge whether the evidence-based requirement of physics justifies the mathematical requirements.
@@
 Additionally, the inner product structure defines the Born rule, the transition probability during measurement.
-While some may consider these physical requirements of quantum mechanics, their status is subject to interpretation.\cite{albert_quantum_1994, wallace_everett_2013, howard_complementarity_2021, ghirardi_unified_1986} Some consider superpositions as mathematical constructs; the Born rule brings in the notion of measurements.
+While some may consider these physical requirements of quantum mechanics, their status is subject to interpretation~\cite{albert_quantum_1994, wallace_everett_2013, howard_complementarity_2021, ghirardi_unified_1986}.
+Some consider superpositions as mathematical constructs; the Born rule brings in the notion of measurements.
 We therefore reframe those requirements in terms of properties of ensembles.
@@
 Since we can physically prepare statistical ensembles, a physical theory must be able to characterize them.
-Mathematically, these form a convex set, where the convex combinations are the statistical mixtures.\footnote{There is a rich literature on generalized probabilistic theories (GPTs) in which people reconstruct quantum theory by imposing requirements on convex spaces.\cite{gpt_overview_2021}} In quantum mechanics, the geometry of the space of statistical ensembles is exactly defined by the complex inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.
+Mathematically, these form a convex set, where the convex combinations are the statistical mixtures.\footnote{There is a rich literature on generalized probabilistic theories (GPTs) in which people reconstruct quantum theory by imposing requirements on convex spaces~\cite{gpt_overview_2021}}.
+In quantum mechanics, the geometry of the space of statistical ensembles is exactly defined by the complex inner product and vector space structures, which means properties of the inner product space can be reexpressed as properties of the space of mixtures.
 For example, we can characterize a superposition in the following way: a pure state $\psi$ is expressible as a superposition of other pure states $\{\phi_i\}$ if and only if there is a mixed state $\rho$ that can be equivalently expressed as a mixture of $\{\phi_i\}$ or as a mixture of $\psi$ and other pure states.
@@
 Mathematically, this means that the space includes the limit of every Cauchy sequence; physically, this property is more difficult to characterize.
-We can look at an equivalent characterization \cite[Theorem 13.8]{roman_2008}.
+We can look at an equivalent characterization~\cite[Theorem 13.8]{roman_2008}.
 An inner product space is complete if, given a series of vectors
@@
 Naively, we could take a region to be any subset of $\mathbb{R}^3$.
-However, this leads to the Banach–Tarski paradox,\cite{tao2021introduction} which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls each of the same volume as the original.
+However, this leads to the Banach–Tarski paradox~\cite{tao2021introduction}, which states that we can take a unit ball, cut it into five pieces, and reassemble those pieces into two balls each of the same volume as the original.
 Clearly, this does not make physical sense.
 However, we get a consistent definition if we restrict regions to be Borel sets, instead of arbitrary subsets.
-Therefore, bigger is not always better in mathematics:\footnote{In another work\cite{aop-book} some of us show that the Borel sets are exactly those sets that can be associated with experimental procedures.
+Therefore, bigger is not always better in mathematics:\footnote{In another work~\cite{aop-book} some of us show that the Borel sets are exactly those sets that can be associated with experimental procedures.
 We have often found that physical requirements naturally map to well-behaved spaces.} we need to look more closely at what the math dragged in.
@@
 On physical grounds, these are clearly not equivalent.
-Reconstructing a state from its expectation values is a common technique, quantum tomography.\cite{banaszek2013focus} If a unitary transformation can change finite expectations to infinite ones, we effectively change what can be distinguished experimentally.
+Reconstructing a state from its expectation values is a common technique, quantum tomography~\cite{banaszek2013focus}.
+If a unitary transformation can change finite expectations to infinite ones, we effectively change what can be distinguished experimentally.
 Infinite uncertainty in one frame would map to finite uncertainty in another.
@@
 We are likely missing out on useful results by not considering what is provable within the subset of physical cases.
-We, therefore, wonder whether many of the technical issues we have in quantum theory (e.g.~the non-existence of a valid measure for path integrals\cite{glimm_quantum_1987}, inequivalent representations for interacting theories\cite{earman_haags_2006}, the non-existence of the categorical tensor product for Hilbert spaces\cite{garrett_tensor_2010, sorkin_inside_2022}, etc.)  are simply a consequence of using unphysical objects at a fundamental level.
+We, therefore, wonder whether many of the technical issues we have in quantum theory (e.g.~the non-existence of a valid measure for path integrals~\cite{glimm_quantum_1987}, inequivalent representations for interacting theories~\cite{earman_haags_2006}, the non-existence of the categorical tensor product for Hilbert spaces~\cite{garrett_tensor_2010, sorkin_inside_2022}, etc.)  are simply a consequence of using unphysical objects at a fundamental level.
 
@@
 
-The Schwartz space has several other interesting properties,\cite{moretti_spectral_2017, reed_methods_1980, hall_quantum_2013, rauch_solution_1991} explored in more detail in the appendix.
+The Schwartz space has several other interesting properties~\cite{moretti_spectral_2017, reed_methods_1980, hall_quantum_2013, rauch_solution_1991}, explored in more detail in the appendix.
 For example:
@@
 \end{itemize}
-These properties make the Schwartz space a much more reasonable candidate to capture the physics.\cite{albert2022bosonic} The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} However, it does not seem to us that this is a final and general answer.
+These properties make the Schwartz space a much more reasonable candidate to capture the physics~\cite{albert2022bosonic}.
+The convergence of expectation values of actual observables replaces convergence over vectors in an abstract space.\footnote{Compare this statement with the ``weak convergence'' of von Neumann algebras.} However, it does not seem to us that this is a final and general answer.
 It is not clear to us how Schwartz spaces are generalized in the case where the topology of the underlying space is not $\mathbb{R}^n$.
@@
 We believe that meaningful progress on the foundations of physics cannot happen without understanding the implicit assumptions embedded in the most basic mathematical tools.
+
+\section*{Acknowledgments}
+The authors wish to thank David Carf\`{i} and Noel Swanson for helpful discussions.
+This paper is part of the ongoing Assumptions of Physics project~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
 
@@
 This appendix presents the same arguments in the article's main body, but expanded with full mathematical detail.
-The figure and the derivations in the Appendix are courtesy of the Assumptions of Physics project.\cite{aop-book}
+The figure and the derivations in the Appendix are courtesy of the Assumptions of Physics project~\cite{aop-book}.
 
@@
 First of all, it is not clear whether superpositions are actual physical entities or whether they are just mathematical constructs.
-This is an area where different interpretations may differ \cite{albert_quantum_1994, wallace_everett_2013, howard_complementarity_2021}, and the physicality of a mathematical representation should not be interpretation-dependent.
+This is an area where different interpretations may differ~\cite{albert_quantum_1994, wallace_everett_2013, howard_complementarity_2021}, and the physicality of a mathematical representation should not be interpretation-dependent.
 Second, in its standard characterization, the Born rule brings in one of the most controversial aspects of the theory, measurements, and thus similarly suffers from an interpretation problem.
@@
 Associating the Born rule with entropy gives us a geometric understanding of the inner product at equal time instead of as a probability of transition, which brings it more in line with the use during superposition.
-It should also be noted that the von Neumann entropy of a mixture can be understood as the minimum Shannon entropy over all possible measurement contexts,\cite{nielsen_chuang_2010} which again shows the connection between the geometry of the convex space of statistical ensembles and the inner product.
+It should also be noted that the von Neumann entropy of a mixture can be understood as the minimum Shannon entropy over all possible measurement contexts~\cite{nielsen_chuang_2010}, which again shows the connection between the geometry of the convex space of statistical ensembles and the inner product.
 
@@
 They were not.
-They are designed to solve another problem, namely extending the spectral theorem to the case of continuous spectra.\cite{Madrid_2005, reed_methods_1980, Landsman2017}
+They are designed to solve another problem, namely extending the spectral theorem to the case of continuous spectra~\cite{Madrid_2005, reed_methods_1980, Landsman2017}.
 
@@
 Physically, this seems too restrictive as it excludes functions like normal distributions.
-At a closer investigation, the Paley-Wiener theorem\cite{fourierCompact} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support.\cite{markushevich2014entire} Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true for distributions in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
+At a closer investigation, the Paley-Wiener theorem~\cite{fourierCompact} tells us that the Fourier transform of a function with compact support is an entire function, and an entire function does not have compact support~\cite{markushevich2014entire}.
+Therefore, we cannot have compact support in both position and momentum.\footnote{This is not true for distributions in classical mechanics, as position and momentum are different variables, not linked by the Fourier transform.
 Therefore the space of distributions with compact support on phase space gives us finite expectation value for all functions of position and momentum.}
@@
 Not only do the expectations of all polynomials exist, but they are enough to identify the elements of the space and define limits.
-The Schwartz space is, in fact, a Fréchet space \cite[Theorem V.9]{reed_methods_1980}, meaning that it is a Hausdorff space, with the topology induced by the seminorms given by the expectations, and it is complete with respect to those seminorms.
+The Schwartz space is, in fact, a Fréchet space~\cite[Theorem V.9]{reed_methods_1980}, meaning that it is a Hausdorff space, with the topology induced by the seminorms given by the expectations, and it is complete with respect to those seminorms.
 Physically, this means that measurements on the polynomials (i.e.~the topology of the seminorms) are enough to identify the states (i.e.~the singletons, which are closed sets as per Hausdorff).\footnote{This essentially tells us that the moment problem restricted to pure states has a unique solution in quantum mechanics.
@@
 If we restrict ourselves to classical pure states, i.e.~points in phase space, then the solution is also unique.
-If we restrict ourselves to compact support, distinct distributions have distinct moments even in the multi-variable case.\cite{moment_problem_2017} If we restrict ourselves to integrable functions, even in the single variable case, distinct distributions can have the same moments.
+If we restrict ourselves to compact support, distinct distributions have distinct moments even in the multi-variable case~\cite{moment_problem_2017}.
+If we restrict ourselves to integrable functions, even in the single variable case, distinct distributions can have the same moments.
 None of these comparisons seems adequate as a quantum state is a statistical object by nature, but it is not an arbitrary mixture.

### 2025-01-29 Proceedings/2025-QPL/CarcassiTalkProposal.tex
Adding QPL talk proposal

\documentclass[10pt,twocolumn, nofootinbib]{revtex4-2}

\usepackage{assumptionsofphysics}
\usepackage{tikz}
\usepackage{breakurl}
\usepackage{tcolorbox}

\newcommand\stcap{\mathrm{scap}}
\newcommand\fraction{\mathrm{frac}}
\newcommand\frcap{\mathrm{fcap}}

\newcommand{\ens}[1][e] {\mathsf{#1}} % Ensemble
\newcommand{\Ens}[1][E] {\mathcal{#1}} % Ensemble space

\newcommand\mix{\mathrm{mix}}
\newcommand\component{\mathrm{comp}}
\newcommand\cospan{\mathrm{cospan}}
\newcommand\dist{\mathrm{dist}}
\newcommand\hull{\mathrm{hull}}
\newcommand\support{\mathrm{supp}}
\newcommand\capacity{\mathrm{scap}}
\newcommand\size{\mathrm{frac}}
\newcommand\fcap{\mathrm{fcap}}

\newcommand\vspan{\mathrm{span}}
\newcommand\cl{\mathrm{cl}}

\def\separate{\downmodels}
\def\nseparate{\ndownmodels}
\def\ortho{\perp}
\def\northo{\nperp}

\def\>{\rangle}
\def\<{\langle}
\DeclareMathOperator{\erf}{erf}

\begin{document}

\title{Ensemble spaces}
\author{Gabriele Carcassi}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}
\author{Christine A. Aidala}
\affiliation{Physics Department, University of Michigan, Ann Arbor, MI 48109}

\date{\today}

\begin{abstract}
	We present the current status of a theory of ensembles that is suitable as a basic theory for all other physical theories. We start with a minimum set of axioms that can be justified from basic physical requirements: experimental verifiability justifies a $T_0$ second countable topology, statistical mixing justifies a convex structure and the variability of ensembles justifies the existence of an entropy function. We show that various results can be recovered from just these three axioms and simple constructions. For example, we can show that an ensembles space must embed in a compact set of a vector space; the entropy induces a metric that generalizes the Fisher-Rao metrics of both classical and quantum mechanics; a more general non-additive probability can be defined which reduces to an additive one in classical mechanics and in quantum measurement context and so on. The goal is to generalize as much as possible the common structures of classical mechanics, quantum mechanics and, hopefully, field theories.

	As the project touches aspects of many different fields (such as topology, function analysis, measure theory, classical and quantum mechanics, information theory, Riemmanian and symplectic geometry), this work is done in consultation with different experts. The main goal of this presentation is to find others that may be interested in this type of integrative work. 
\end{abstract}

\maketitle

\section{Introduction}

The goal of our work is to find a minimal set of physically justifiable axioms from which we can build a theory that can serve as a basis for all physical theories. This seems to resonate with people working on reconstructions of quantum mechanics, in particular those working on Generalized Probability Theories. It was the recommendation of some of these to submit to QPL.

The main difference in our approach is that we start at a more basic level and try to incorporate results from different areas (e.g. non-additive measure theory, Riemmanian and symplectic geometry, variational calculus, classical and quantum information theory, function analysis). That is, we want to find a minimal physically justifiable starting points from which most of these structure can be recovered with tighter mathematical constraints.

As the project is, by its nature, rather interdisciplinary, and the publishing world is not setup to encourage the proper review of manuscript that cuts across so many field, the pre-print is effectively a new chapter of our Assumptions of Physics. Openly available, we constantly review the results with experts of different field, gaining a lot more feedback than what peer review allows. We enclose a snapshot of the chapter, though a more advanced version is available online. This is the only way we have found so that we can have every part reviewed by experts from different fields, while making sure all pieces fits together.

\section{Basic requirements for a physical theory}

The strategy is to find requirements that are constitutive of a physical theory. That is, we want to argue from requirement that are necessary to define a physical system and a physical theory, so that the only choice is between accepting those requirement or do nothing at all.

In past work we argued that since science is about statement that are connected to experimental evidence, they require the notion of a verifiable statement: a statement that has an associated test that terminates successfully in finite time if and only if the statement is true. We should that the requirement of finite time verifiability leads to a structure where the experimentally defined possible cases are points of a $T_0$ \textbf{second countable topological space}, where the open sets are verifiable statements and Borel sets are statement associated to a test, regardless of termination. This gives a solid fundamental basis, that is able to justified even details like the analytical discontinuity of quantities over phase transitions.

We now argue that, since the goal of physics is to find general laws that are repeatably testable, the subject of these laws must be ensembles. Laws, in fact, are relationships of the type ``whenever we have this, we have that'' and therefore relate one collection of instances to another collection of instances. Repeatability requires to be able to carry out the tests always ``one more time,'' which implicitly requires the instances to be infinite. Moreover, experimentally our preparation procedures are never exact, so both preparations and measurement are in terms of statistical objects. Therefore, any physical theory must \textbf{at least} define what ensembles the theory allows. Having justified the existence of an ensemble space in any physical theory, we not want to formulate a minimal set of axioms for these spaces.

\section{Ensemble spaces}

Ensembles must have, at least, the following three features: they have to be identifiable experimentally; they must allow statistical mixtures; their variability, their entropy, must be well-defined. Classical and quantum theory do satisfy these axioms. Let us go through these briefly, leaving the full justification to the supporting material.

\subsection{Experimental verifiability}
The first requirement is that ensembles must be connected to experimental verification. That is, we must have enough \textbf{verifiable statements} at our disposal to define ensembles and tell them apart. As mentioned before, we have already developed a full theory for this part, therefore this requirement justifies a $T_0$ second-countable topology on the space of ensemble $\Ens$. It is crucial to understand that the open sets represent the only statements experimentally verifiable, as this will put additional consistency requirements. For example, it would make no sense to assign non-zero probability to a set of cases that can never be experimentally reached.

\begin{tcolorbox}[colback=white, colframe=black]
	Experimental verifiability $\Rightarrow$ An ensemble space $\Ens$ is a $T_0$ second-countable topological space.
\end{tcolorbox}

\subsection{Statistical mixtures}
Another basic requirement is the ability to prepare a \textbf{mixture} of two ensembles. If $\ens[a], \ens[b] \in \Ens$ are two ensembles, $p \in [0,1]$ a weight and $\bar{p} = 1-p$, then $\ens = p \ens[a] + \bar{p} \ens[b] \in \Ens$ is the ensemble that describes a process that selects the first ensemble over the second $p$ percent of the times. Mathematically, the ensemble space is endowed with a \textbf{convex structure}. Moreover, since the mixing operation must be consistent with experimental verifiability, it will be topologically continuous, and the ensemble space will be a topological convex space.

\begin{tcolorbox}[colback=white, colframe=black]
	Statistical mixture of ensembles $\Rightarrow$ An ensemble space $\Ens$ is a topological convex space.
\end{tcolorbox}

\subsection{Entropy}
The last requirement is that, since each ensemble represents an infinite collection of preparations, the variability among those preparation is well-defined. That is, we have a well-defined \textbf{entropy} that quantifies us how much the elements in each ensemble are different from each other.

The variability will have consistency requirement with respect to mixing. Intuitively, the variability cannot decrease through a statistical mixture, and the maximum increase is given when to ensembles are \textbf{orthogonal} (noted $\ens[a] \ortho \ens[b]$), which means no instance from the first ensemble could have been prepared from the second and vice-versa. The entropy $S : \Ens \to \mathbb{R}$, then, satisfies the following conditions:
\begin{enumerate}
	\item strictly concave: $S(p \ens[a] + \bar{p} \ens[b]) - (p S(\ens[a]) + \bar{p} S(\ens[b]) ) \geq 0$
	\item upper variability bound: $S(p \ens[a] + \bar{p} \ens[b]) - (p S(\ens[a]) + \bar{p} S(\ens[b]) ) \leq I(p,\bar{p})$; we define $\ens[a] \ortho \ens[b]$ if equality holds
	\item mixtures preserves orthogonality: $\ens \ortho p \ens[a] + \bar{p} \ens[b]$ if and only if $\ens \ortho \ens[a]$ and $\ens \ortho \ens[b]$.
\end{enumerate}
Note that the function $I$ is left unspecified. These conditions are enough to show that $I$ must be the Shannon entropy, up to a multiplicative constant, which is now a general result independent of what type of ensemble space one has.

\begin{tcolorbox}[colback=white, colframe=black]
	Existence of entropy $\Rightarrow$ Strictly concave function whose upper bound defines orthogonality.
\end{tcolorbox}

\section{Results and open questions}

The rest of the presentation is a series of results that can be derived on these three axioms alone.

\textbf{Separateness vs orthogonality.} We can define two notions of ``distinctness'' among ensembles. Two ensembles are separate, $\ens[a] \separate \ens[b]$, if, intuitively, they have no common component. That is, there is no $\ens[c]$, the common component, such that $\ens[a] = p_1 \ens[c] + \bar{p}_1 \ens_1$ and $\ens[b] = p_2 \ens[c] + \bar{p}_2 \ens_2$. Two ensembles are orthogonal, $\ens[a] \ortho \ens[b]$, if they saturate the upper variability bound. One can prove that orthogonality implies separateness, but the reverse is not true.

In classical spaces, the two coincides, which already gives a simple characterization for classical spaces: two ensembles that have no common component saturate the upper entropy bound. A lot of structure comes out between the interplay of these two simple concepts.

\textbf{Entropic constraints on the convex space.} Imposing a convex structure only leaves over the door for pathological cases that the existence of the entropy excludes. The entropy, in fact, forces the convex space to embed into a vector space and each affine line will intersect the ensemble space for a finite interval. Specifically, it is the bounds on the entropy that imposes these constraints.

Whether it embeds continuously into a topological vector space it still needs to be determined.

\textbf{Convex sets and flats.} Note that the mixing operation can only be extended to finite convex combinations. This is intended as there is no mathematically safe and physically meaningful way to close under infinite combinations. For example, a probability distribution over countably many elements can have an infinite Shannon entropy. It is the topology that determines whether a specific infinite mixture is allowed or not. For this reason, we will always consider convex sets to be closed both under the convex structure and the topology. The $\hull$ of a set $A \subset \Ens$ is the closure on both structures, which represents all mixtures, finite or infinite, that can be generated from $A$.

A flat, instead, will be closed on affine mixtures as well. While convex combinations always go ``inside'' the space, affine mixtures will extends the subspace to the boundaries of the ensemble space. Different structures will require the different closures, leading to a richer structure.

\textbf{Entropic geometry.} The entropy imposes, without extra conditions, a geometric structure. We can define the mixing entropy $MS(\ens[a], \ens[b]) = S\left(\frac{1}{2}\ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2} S(\ens[a]) + \frac{1}{2} S(\ens[b])\right).$ Conceptually, this tells us how much the entropy increases when mixing two elements. This quantity is non-negative, it is zero if and only if $\ens[a] = \ens[b]$ and it is one if and only if $\ens[a] \ortho \ens[b]$. While it does not satisfy the triangle inequality, this pseudo-distance tells us how ``different'' two ensembles are. It recovers the Jensen-Shannon divergence in both the classical and the quantum case.

Since the mixing entropy is continuous, it can be used to generate open balls that are open sets in the topology. If it can be proved that, given an open set, every point has an open ball that is fully contained in the set, this would show that the topology is the one generated by entropy in such a way. This is still an open question. Moreover, in both classical and quantum mechanics, the Jensen-Shannon divergence is the square of a distance function. It is unknown whether this extends to the general case.

Since the entropy is strictly concave, its Hessian is negative definite. Its negation, then, is a positive-definite bilinear form of variations. This can be understood as a metric tensor. In classical and quantum mechanics, this recovers the Fisher-Rao information metric. Some detail still need to be nailed down to make sure the construction is mathematically well-defined.

\textbf{Fraction capacity and non-additive probability.} Given a target ensemble $\ens$ we can ask what part can be described by another ensemble $\ens[a]$. We define the \textbf{fraction} of $\ens[a]$ in $\ens$ to be $\fraction_{\ens}(\ens[a]) = \sup(\{ p \in [0,1] \, | \, \exists \, \ens_1 \in \Ens \text{ s.t. }  \ens = p \ens[a] + \bar{p} \ens_1 \})$. Given a set $A$, instead, we can ask what is the biggest component of $\ens$ that can be expressed as a convex combination of $A$. We define the\textbf{fraction capacity} of $A$ for $\ens$ to be $\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\})$.

The fraction capacity is then a set function that is non-negative, unit bounded, monotone, sub-additive, continuous from below and above and additive over orthogonal sets. It can be understood as a non-additive generalization of a probability measure that can only be defined on any physically meaningful state space.

\textbf{Classical probability contexts.} We then define a classical probability context as a flat that does not allow multiple decomposition. In the finite dimensional case, this reduces to a simplex. Every classical space is also a classical probability context. The space of density operators that commute with a maximal set of observable also form a classical probability context. In these cases, we have developed a construction that allows us to show that each ensemble in a classical probability context is a classical probability measure.

The construction allows us to create the points (e.g. the spectrum of possible values of position and momentum in classical mechanics) from the ensembles themselves. That is, given an ensemble space we do not know a priory whether the elements are distributions, and the points of the space are constructed as limits of ensembles that overlap less and less. Since the construction is physically motivated and requires no extra assumptions, the hope is that it can be used in field theories and a possible theory of quantum gravity.

\textbf{State capacity.} Given an ensemble, we can ask what is its spread over physically distinguishable cases. It can be argued that the spread must be a monotonic function of the entropy, and must be subadditive. The exponential of the entropy has these properties, and can be used as a measure of volume of states. Given a set, we can define the state capacity as $\capacity(A) = \sup(2^{S(\hull(A))}\cup\{0\})$. This is another set function that is non-negative, monotone, subadditive and additive over orthogonal sets.

The idea would be to use both the fraction and state capacity to extend the measure theoretic concepts to the non-additive case, so that we can create a non-additive theory of probability that can work for all physical theories.

\textbf{Conclusion.} We found that these results were interesting to people working on foundations, particularly working on Generalized Probability Theories.

\section*{Acknowledgments}
This paper is part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived. This work was made possible through the support of grant \#62847 from the John Templeton Foundation.

\bibliography{bibliography}

\newcommand{\pj}[1] {\underbar{$#1$}}

\end{document}

### 2025-02-04 Proceedings/2025-QPL/CarcassiTalkProposal.tex
QPL2025: submitted talk proposal

@@
 
+\usepackage{hyperref}
+\hypersetup{
+	colorlinks=true,
+	urlcolor=blue,
+	linkcolor=blue,
+	citecolor=black
+}
+\urlstyle{same}
+
 \begin{document}
@@
 	We present the current status of a theory of ensembles that is suitable as a basic theory for all other physical theories.
-We start with a minimum set of axioms that can be justified from basic physical requirements: experimental verifiability justifies a $T_0$ second countable topology, statistical mixing justifies a convex structure and the variability of ensembles justifies the existence of an entropy function.
+We start with a minimum set of axioms that can be justified from basic physical requirements: experimental verifiability justifies a $T_0$ second-countable topology, statistical mixing justifies a convex structure and the variability of ensembles justifies the existence of an entropy function.
 We show that various results can be recovered from just these three axioms and simple constructions.
-For example, we can show that an ensembles space must embed in a compact set of a vector space; the entropy induces a metric that generalizes the Fisher-Rao metrics of both classical and quantum mechanics; a more general non-additive probability can be defined which reduces to an additive one in classical mechanics and in quantum measurement context and so on.
+For example, we can show that an ensemble space must embed in a compact set of a vector space; the entropy induces a metric that generalizes the Fisher-Rao metrics of both classical and quantum mechanics; a more general non-additive probability can be defined which reduces to an additive one in classical mechanics and in quantum measurement context and so on.
 The goal is to generalize as much as possible the common structures of classical mechanics, quantum mechanics and, hopefully, field theories.
 
-	As the project touches aspects of many different fields (such as topology, function analysis, measure theory, classical and quantum mechanics, information theory, Riemmanian and symplectic geometry), this work is done in consultation with different experts.
+	As the project touches aspects of many different fields (such as topology, function analysis, measure theory, classical and quantum mechanics, information theory, Riemannian and symplectic geometry), this work is done in consultation with different experts.
 The main goal of this presentation is to find others that may be interested in this type of integrative work.
@@
 This seems to resonate with people working on reconstructions of quantum mechanics, in particular those working on Generalized Probability Theories.
-It was the recommendation of some of these to submit to QPL.
+It was the recommendation of some of these researchers to submit to QPL.
 
 The main difference in our approach is that we start at a more basic level and try to incorporate results from different areas (e.g.
-non-additive measure theory, Riemmanian and symplectic geometry, variational calculus, classical and quantum information theory, function analysis).
-That is, we want to find a minimal physically justifiable starting points from which most of these structure can be recovered with tighter mathematical constraints.
+non-additive measure theory, Riemannian and symplectic geometry, variational calculus, classical and quantum information theory, functional analysis).
+That is, we want to find a set of minimal physically justifiable starting points from which most of these structures can be recovered with tighter mathematical constraints.
 
-As the project is, by its nature, rather interdisciplinary, and the publishing world is not setup to encourage the proper review of manuscript that cuts across so many field, the pre-print is effectively a new chapter of our Assumptions of Physics.
-Openly available, we constantly review the results with experts of different field, gaining a lot more feedback than what peer review allows.
-We enclose a snapshot of the chapter, though a more advanced version is available online.
-This is the only way we have found so that we can have every part reviewed by experts from different fields, while making sure all pieces fits together.
+As the project is, by its nature, rather interdisciplinary, and the publishing world is not set up to encourage the proper review of manuscripts that cut across so many field, the preprint is effectively a new chapter of our book \textit{Assumptions of Physics}~\cite{aop-book}.
+Openly available, we constantly review the results with experts of different fields, gaining a lot more feedback than what peer review allows.
+We enclose a snapshot of the chapter, though a more advanced version is  \href{https://latexonline.cc/compile?git=https%3A%2F%2Fgithub.com%2Fassumptionsofphysics%2Fbook&target=art_pm_EnsembleSpaces.tex&command=pdflatex}{available online}.
+This is the only way we have found to have every part reviewed by experts from different fields, while making sure all pieces fit together.
 
@@
 The strategy is to find requirements that are constitutive of a physical theory.
-That is, we want to argue from requirement that are necessary to define a physical system and a physical theory, so that the only choice is between accepting those requirement or do nothing at all.
+That is, we want to argue from requirements that are necessary to define a physical system and a physical theory, so that the only choice is between accepting those requirements or doing nothing at all (e.g.~if one does not accept the existence of testable statements, one cannot do science).
 
-In past work we argued that since science is about statement that are connected to experimental evidence, they require the notion of a verifiable statement: a statement that has an associated test that terminates successfully in finite time if and only if the statement is true.
-We should that the requirement of finite time verifiability leads to a structure where the experimentally defined possible cases are points of a $T_0$ \textbf{second countable topological space}, where the open sets are verifiable statements and Borel sets are statement associated to a test, regardless of termination.
-This gives a solid fundamental basis, that is able to justified even details like the analytical discontinuity of quantities over phase transitions.
+In past work we argued that since science is about statements that are connected to experimental evidence, they require the notion of a verifiable statement: a statement that has an associated test that terminates successfully in finite time if and only if the statement is true.
+We showed~\cite{aop-book,aop-math-topologydistinguishability} that the requirement of finite-time verifiability leads to a structure where the experimentally defined possible cases are points of a $T_0$ \textbf{second-countable topological space}, where the open sets are verifiable statements and Borel sets are statements associated to a test, regardless of termination.
+This gives a solid fundamental basis, which is able to justify even details like the analytical discontinuity of quantities over phase transitions.
 
 We now argue that, since the goal of physics is to find general laws that are repeatably testable, the subject of these laws must be ensembles.
-Laws, in fact, are relationships of the type ``whenever we have this, we have that'' and therefore relate one collection of instances to another collection of instances.
-Repeatability requires to be able to carry out the tests always ``one more time,'' which implicitly requires the instances to be infinite.
-Moreover, experimentally our preparation procedures are never exact, so both preparations and measurement are in terms of statistical objects.
+Physical laws, in fact, are relationships of the type ``whenever we have this, we have that'' and therefore relate one collection of instances to another collection of instances.
+Repeatability requires the ability to carry out the tests always ``one more time,'' which implicitly requires the instances to be infinite.
+Moreover, experimentally our preparation procedures are never exact, so both preparations and measurements are in terms of statistical objects.
 Therefore, any physical theory must \textbf{at least} define what ensembles the theory allows.
-Having justified the existence of an ensemble space in any physical theory, we not want to formulate a minimal set of axioms for these spaces.
+Having justified the existence of an ensemble space in any physical theory, we now want to formulate a minimal set of axioms for these spaces.
 
@@
 
-Ensembles must have, at least, the following three features: they have to be identifiable experimentally; they must allow statistical mixtures; their variability, their entropy, must be well-defined.
-Classical and quantum theory do satisfy these axioms.
+Ensembles must have, at least, the following three features: they have to be identifiable experimentally; they must allow statistical mixtures; their variability, their entropy, must be well defined.
+Classical and quantum theory both satisfy these axioms.
 Let us go through these briefly, leaving the full justification to the supporting material.
@@
 That is, we must have enough \textbf{verifiable statements} at our disposal to define ensembles and tell them apart.
-As mentioned before, we have already developed a full theory for this part, therefore this requirement justifies a $T_0$ second-countable topology on the space of ensemble $\Ens$.
-It is crucial to understand that the open sets represent the only statements experimentally verifiable, as this will put additional consistency requirements.
+As mentioned before, we have already developed a full theory for this part, therefore this requirement justifies a $T_0$ second-countable topology on the space of ensembles $\Ens$.
+It is crucial to understand that the open sets represent the only statements that are experimentally verifiable, as this will impose additional consistency requirements.
 For example, it would make no sense to assign non-zero probability to a set of cases that can never be experimentally reached.
@@
 Another basic requirement is the ability to prepare a \textbf{mixture} of two ensembles.
-If $\ens[a], \ens[b] \in \Ens$ are two ensembles, $p \in [0,1]$ a weight and $\bar{p} = 1-p$, then $\ens = p \ens[a] + \bar{p} \ens[b] \in \Ens$ is the ensemble that describes a process that selects the first ensemble over the second $p$ percent of the times.
+If $\ens[a], \ens[b] \in \Ens$ are two ensembles, $p \in [0,1]$ a weight and $\bar{p} = 1-p$, then $\ens = p \ens[a] + \bar{p} \ens[b] \in \Ens$ is the ensemble that describes a process that selects the first ensemble over the second a fraction $p$ of the times.
 Mathematically, the ensemble space is endowed with a \textbf{convex structure}.
@@
 \subsection{Entropy}
-The last requirement is that, since each ensemble represents an infinite collection of preparations, the variability among those preparation is well-defined.
-That is, we have a well-defined \textbf{entropy} that quantifies us how much the elements in each ensemble are different from each other.
+The last requirement is that, since each ensemble represents an infinite collection of preparations, the variability among those preparations is well defined.
+That is, we have a well-defined \textbf{entropy} that quantifies the extent to which the elements in each ensemble are different from each other.
 
-The variability will have consistency requirement with respect to mixing.
-Intuitively, the variability cannot decrease through a statistical mixture, and the maximum increase is given when to ensembles are \textbf{orthogonal} (noted $\ens[a] \ortho \ens[b]$), which means no instance from the first ensemble could have been prepared from the second and vice-versa.
+The variability will have consistency requirements with respect to mixing.
+Intuitively, the variability cannot decrease through a statistical mixture, and the maximum increase is given when two ensembles are \textbf{orthogonal} (noted $\ens[a] \ortho \ens[b]$), which means no instance from the first ensemble could have been prepared from the second and vice-versa.
 The entropy $S : \Ens \to \mathbb{R}$, then, satisfies the following conditions:
@@
 	\item upper variability bound: $S(p \ens[a] + \bar{p} \ens[b]) - (p S(\ens[a]) + \bar{p} S(\ens[b]) ) \leq I(p,\bar{p})$; we define $\ens[a] \ortho \ens[b]$ if equality holds
-	\item mixtures preserves orthogonality: $\ens \ortho p \ens[a] + \bar{p} \ens[b]$ if and only if $\ens \ortho \ens[a]$ and $\ens \ortho \ens[b]$.
+	\item mixtures preserve orthogonality: $\ens \ortho p \ens[a] + \bar{p} \ens[b]$ if and only if $\ens \ortho \ens[a]$ and $\ens \ortho \ens[b]$.
 \end{enumerate}
@@
 
-The rest of the presentation is a series of results that can be derived on these three axioms alone.
+The rest of the presentation is a series of results that can be derived from these three axioms alone.
 
-\textbf{Separateness vs orthogonality.} We can define two notions of ``distinctness'' among ensembles.
+\textbf{Separateness vs.~orthogonality.} We can define two notions of ``distinctness'' among ensembles.
 Two ensembles are separate, $\ens[a] \separate \ens[b]$, if, intuitively, they have no common component.
@@
 Two ensembles are orthogonal, $\ens[a] \ortho \ens[b]$, if they saturate the upper variability bound.
-One can prove that orthogonality implies separateness, but the reverse is not true.
+One can prove that orthogonality implies separateness, but the converse is not true.
 
-In classical spaces, the two coincides, which already gives a simple characterization for classical spaces: two ensembles that have no common component saturate the upper entropy bound.
+In classical spaces, the two coincide, which already gives a simple characterization for classical spaces: two ensembles that have no common component saturate the upper entropy bound.
 A lot of structure comes out between the interplay of these two simple concepts.
 
-\textbf{Entropic constraints on the convex space.} Imposing a convex structure only leaves over the door for pathological cases that the existence of the entropy excludes.
+\textbf{Entropic constraints on the convex space.} Imposing only a convex structure leaves open the door for pathological cases that the existence of the entropy excludes.
 The entropy, in fact, forces the convex space to embed into a vector space and each affine line will intersect the ensemble space for a finite interval.
-Specifically, it is the bounds on the entropy that imposes these constraints.
+Specifically, it is the bounds on the entropy that impose these constraints.
 
-Whether it embeds continuously into a topological vector space it still needs to be determined.
+Whether it embeds continuously into a topological vector space still needs to be determined.
 
-\textbf{Convex sets and flats.} Note that the mixing operation can only be extended to finite convex combinations.
+\textbf{Convex sets and flats.} Note that the mixing operation can only be guaranteed up to finite convex combinations.
 This is intended as there is no mathematically safe and physically meaningful way to close under infinite combinations.
@@
 For this reason, we will always consider convex sets to be closed both under the convex structure and the topology.
-The $\hull$ of a set $A \subset \Ens$ is the closure on both structures, which represents all mixtures, finite or infinite, that can be generated from $A$.
+The $\hull$ of a set $A \subseteq \Ens$ is the closure on both structures, which represents all mixtures, finite or infinite, that can be generated from $A$.
 
 A flat, instead, will be closed on affine mixtures as well.
-While convex combinations always go ``inside'' the space, affine mixtures will extends the subspace to the boundaries of the ensemble space.
+While convex combinations always stay ``inside'' the subspace, affine mixtures extend the subspace to the boundaries of the ensemble space.
 Different structures will require the different closures, leading to a richer structure.
@@
 In classical and quantum mechanics, this recovers the Fisher-Rao information metric.
-Some detail still need to be nailed down to make sure the construction is mathematically well-defined.
+Some details still need to be nailed down to make sure the construction is mathematically well-defined.
 
@@
 }  \ens = p \ens[a] + \bar{p} \ens_1 \})$.
-Given a set $A$, instead, we can ask what is the biggest component of $\ens$ that can be expressed as a convex combination of $A$.
-We define the\textbf{fraction capacity} of $A$ for $\ens$ to be $\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\})$.
+Given a set $A$, instead, we can ask what is the biggest component of $\ens$ that can be expressed as a convex combination of elements of $A$.
+We define the \textbf{fraction capacity} of $A$ for $\ens$ to be $\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\})$.
 
-The fraction capacity is then a set function that is non-negative, unit bounded, monotone, sub-additive, continuous from below and above and additive over orthogonal sets.
+The fraction capacity is then a set function that is non-negative, unit bounded, monotone, subadditive, continuous from below and above and additive over orthogonal sets.
 It can be understood as a non-additive generalization of a probability measure that can only be defined on any physically meaningful state space.
@@
 \textbf{Classical probability contexts.} We then define a classical probability context as a flat that does not allow multiple decomposition.
-In the finite dimensional case, this reduces to a simplex.
+In the finite-dimensional case, this reduces to a simplex.
 Every classical space is also a classical probability context.
-The space of density operators that commute with a maximal set of observable also form a classical probability context.
+The space of density operators that commute with a maximal set of observables also form a classical probability context.
 In these cases, we have developed a construction that allows us to show that each ensemble in a classical probability context is a classical probability measure.
@@
 the spectrum of possible values of position and momentum in classical mechanics) from the ensembles themselves.
-That is, given an ensemble space we do not know a priory whether the elements are distributions, and the points of the space are constructed as limits of ensembles that overlap less and less.
-Since the construction is physically motivated and requires no extra assumptions, the hope is that it can be used in field theories and a possible theory of quantum gravity.
+That is, given an ensemble space we do not know \textit{a priori} whether the elements are distributions, and the points of the space are constructed as limits of ensembles that overlap less and less.
+Since the construction is physically motivated and requires no extra assumptions, the hope is that it can be used in field theories and in a possible theory of quantum gravity.
 
-\textbf{State capacity.} Given an ensemble, we can ask what is its spread over physically distinguishable cases.
+\textbf{State capacity.} Given an ensemble, we can ask what its spread is over physically distinguishable cases.
 It can be argued that the spread must be a monotonic function of the entropy, and must be subadditive.
@@
 
-The idea would be to use both the fraction and state capacity to extend the measure theoretic concepts to the non-additive case, so that we can create a non-additive theory of probability that can work for all physical theories.
-
-\textbf{Conclusion.} We found that these results were interesting to people working on foundations, particularly working on Generalized Probability Theories.
+The idea would be to use both the fraction and state capacity to extend the measure-theoretic concepts to the non-additive case, so that we can create a non-additive theory of probability that can work for all physical theories.
 
@@
 This paper is part of the ongoing \textit{Assumptions of Physics} project \cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
-This work was made possible through the support of grant \#62847 from the John Templeton Foundation.
+This work was made possible in part through the support of grant \#62847 from the John Templeton Foundation.
 

### 2025-02-24 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: resumed working on the article

@@
 
+\def\ortho{\perp}
+\def\northo{\nperp}
+\def\separate{\downmodels}
+\def\nseparate{\ndownmodels}
+
 \def\>{\rangle}
 \def\<{\langle}
-\DeclareMathOperator{\erf}{erf}
 
@@
 fuzzy) measure theory.
-We will define a generalized concept of probability that is ultimately based on physically motivated axioms, giving the physicist a clear operational motivation to our definitions, the mathematicians a formally well-defined mathematical structure and the philosopher a unified framework in which probability can be interpreted in the context of physical theories.
+We will define a generalized concept of probability that is ultimately based on physically motivated axioms, giving the physicist a clear operational motivation to our definitions, the mathematicians a formally well-defined mathematical structure and the philosopher a unified framework in which probability can be interpreted in the context of all physical theories.
 We are not going to be able to develop a full theory here, as there are still a number of conceptual and technical challenges to be solved that will require work in physics, mathematics and philosophy.
-However, we will show that there are enough results that point in a good direction, and therefore make it reasonable to invest time and effort to work on the open problems.
-For those interested, the current state of the research will always be available in an open document that will be routinely updated as the theory matures.
+However, we will show that there are enough results that point in the same direction, and therefore make it reasonable to invest time and effort to work on the open problems.
+
+The material presented is part of a larger work that aims to coalesce ideas and results from different branch of mathematics and physics.
+For those interested, the current state of the research is openly available in an open document that will be routinely updated as experts from different fields review and help us improve the overall theory.
 
@@
 
+Similarly, given a set $A$ of ensembles we define the \textbf{state capacity} as the exponential of the highest entropy obtainable with a mixture of $A$.
+Note that the highest entropy of a subspace in both classical and quantum statistical mechanics is given by the uniform distribution, therefore the state capacity recovers the count of states (i.e.
+the Liouville volume) in classical mechanics and the dimensionality of the Hilbert space in quantum mechanics.
+The state capacity is also a non-negative, sub-additive and continuous set function.
+This points to a possible generalization of measure theoretic calculus to a sub-additive case that would be guaranteed to work on all physical theories.
+
 \section{Quick review of probability theory}
+
+The standard way to characterize probability is through measure theory.
+Given a sample space $\Omega$, which represents the possible outcomes, and a $\sigma$-algebra $\Sigma_{\Omega}$, which represents the events, we define a probability measure $p : \Sigma_{\Omega} \to [0,1]$.
+Note that the probability is assigned to events, not the outcomes.
+This is what allows to treat discrete and continuous spaces within the same framework.
+In physics, all sample spaces are ultimately topological spaces, and the $\sigma$-algebra is typically the Borel algebra (i.e.
+the smallest $\sigma$-algebra that contains all open sets).
+
+In classical mechanics, the sample space is phase space $M$ given by all possible values of position and momentum for all particles under considerations.
+Phase space is equipped with another measure, the Liouville measure $\mu : \Sigma_M \to [0, +\infty]$, that returns volume of each region.
+In statistical mechanics, this volume represents the count of states and is another key ingredient.
+The probability density over states is given by the Radon-Nikodym derivative $\rho = \frac{dp}{d\mu}$ and the entropy is given by $S(\rho) = - \int_M \rho \log \rho d\mu$.
+In case of a uniform distribution $\rho_U$ with support $U$, the entropy is given by $S(\rho_U) = \log \mu(U)$.\footnote{Note that entropy, the count of states and the symplectic form that characterizes phase space are deeply intertwined, yet they require different mathematical framework to express them.
+We suspect that they are "shadows" of a more fundamental mathematical structure, which our work on ensembles space is trying to identify.}
+
+The above definitions do not apply to quantum mechanics.
+In the Hilbert space formulation, mixed states are represented by density operator (i.e.
+positive semi-definite self-adjoint operators with trace one).
+Note that if $\rho(x)$ is an integrable function such that $\int_X \rho(x) dx = 1$, the map $\psi(x) \mapsto \rho(x) \psi(x)$ is a density operator.
+Therefore the space of probability densities over position are mixed states in quantum mechanics.
+However, no joint distribution between position and momentum can be given, therefore quantum states cannot be described by the space of probability measures over a single sample space.
+
+Alternative representations over phase space do exist, and they are widely used.(CITE) The use quasi-probability measure, meaning that the measure over the full space is still unitary (i.e.
+$\mu(M) =1$).
+However, the measure can be negative for smaller regions, which means it can also be grater than $1$, and there is no simple characterization of which quasi-probability measures are allowed and which aren't.
+Moreover, there is no simple link between then entropy of the distribution and the Liouville measure.
 
@@
 
-As mentioned in the introduction, our goal is to construct a probability theory of general applicability to physical systems.
-Our approach is not to define an axiomatic theory of probability and apply it to physical systems, but rather to characterize basic properties of physical states and from those build a theory of probability that fits those definitions.
+Note that our goal is not to construct an abstract probability theory that applies to everything, but rather to construct a specific probability theory that is guaranteed to apply to all physical systems.
+Therefore we do not to define an axiomatic theory of probability and apply it to physical systems, but rather characterize basic properties of physical states and from those build a theory of probability that fits those definitions.
+Whether it is applicable to decision theory, economics, or psychology is out of scope.
 As is common in our larger project, we start from axioms that, in our opinion, represent requirements of scientific practice.
+We call these \textbf{constitutive assumptions} as they are essential to proceed with the scientific endeavor.
 The advantage is that the realm of applicability and physical interpretation will be clear from the start.
 
-We start from requiring a physical theory to allow statistical descriptions: at the very least, it will provide us with a notion of ensemble, and the space of ensembles that are allowed by the theory.
+We start from requiring a physical theory to allow statistical descriptions: at the very least, it will provide a notion of ensemble and describe which ensembles that are allowed by the theory.
 An ensemble can be understood as the infinite collection of the outputs of a particular preparation procedure.
-The ensemble space of classical mechanics, for example, is the space of continuous probability distributions on phase space.
+The ensemble space of classical mechanics, for example, is the space of probability distributions on phase space.
 In quantum mechanics, instead, it is the space of density operators.
@@
 Second, if a physical theory is to be testable experimentally, and repeatedly so, it must be in terms of ensembles because this is what we prepare in practice and this is what, in the end, we characterize experimentally.
-The approach, then, can most likely be given firm conceptual grounds, which we leave as an open but solvable philosophical problem.
+Moreover, repeatability requires the ability to always test ``one more time,'' which implies that ensembles are infinite collections, and therefore idealization.
+The approach, then, can most likely be given firm conceptual grounds, which we leave as an open but likely solvable philosophical problem.
 
@@
 For example, the statement ``the mass of the photon is less than $10^{-13} \, eV$'' is verifiable, while ``the mass of the photon is exactly $0 \, eV$'' is not verifiable due to its infinite precision.
-Based on previous work, this mathematically requires an ensemble space to be a $T_0$ second-countable topological space, where open sets represent verifiable statements and Borel sets represent statements associated with a test, but without guarantee of termination.
+Based on previous work, (CITE) this mathematically requires an ensemble space to be a $T_0$ second-countable topological space, where open sets represent verifiable statements and Borel sets represent statements associated with a test, but without guarantee of termination.
 
@@
 Ultimately, the convex structure is responsible for all linear structures we have in physics.
-Both in classical and quantum mechanics, the ensemble space is not just a convex set, but a convex subset of a vector space.
-In those cases, in fact, the convex structure is, in a sense, invertible.
-If we fix $\ens$, $\ens[a]$ and $p$ then, if it exists, there is only one $\ens[b]$ such that $\ens = p \ens[a] + (1-p) \ens[b]$.
-The ensemble space is \textbf{complemented}, meaning that there is only one complement to a component of an ensemble.\footnote{In the context of convex spaces, this property is called ``cancellative.''}
+In face, the entropy and its bounds, defined later, forces the ensemble space to be the subset of a convex subset of a vector space.
+That is, the convex structure is ``invertible.''\footnote{If we fix $\ens$, $\ens[a]$ and $p$ then, if it exists, there is only one $\ens[b]$ such that $\ens = p \ens[a] + (1-p) \ens[b]$.
+In the context of convex spaces, this property is called ``cancellative.''} What is still not clear is whether the ensemble space embeds continuously in a topological vector space.
+While topological vector spaces are well established in the literature, topological convex spaces are not.
 
-This perspective allows us to understand what negative probability is.
+Regardless of the topology, we are already equipped to understand what negative probability is.
 Suppose we have a real vector space of dimensions $n$, then $n$ linearly independent vectors are enough to express every point as a linear combination.
-If we add another point, we can now express every point as an affine combination, that is a linear combination where the coefficients sum to one, though some can still be negative.
-Since the space of mixed states in quantum mechanics is a real vector space, we can choose a linearly dependent set of states that span the whole space, and every mixed state can be expressed as an affine combination, as a pseudo-probability.
+We can also take an additional vector and express every point as an affine combination, that is a linear combination where the coefficients sum to one, though some can still be negative.
+Since every ensemble space, including the space of mixed states in quantum mechanics, is a subset of a real vector space, we can choose a linearly dependent set of states that span the whole space, and every mixed state can be expressed as an affine combination, as a pseudo-probability.
+However, the boundaries of the ensemble space will determine which affine combinations are valid states (i.e.
+they end up within the ensemble space), and there is no requirement that these boundary are conveniently expressible in terms of affine combinations.
 
-Pseudo-probability distributions, then, require the additional vector space structure.
-However, we found no physical justification to require the additional structure, and we suspect that it may need to be relaxed in future physical theories.
-Therefore it is important that our fundamental definitions require only the convex structure, and not the vector space structure.
+The convex structure, however, allows us to ask whether two ensembles have a \textbf{common component}.
+That is, then can be seen as different statistical mixture that have some part in common.
+Two ensemble are separate, noted $\ens[a] \separate \ens[b]$, if they do not have a common component.
+This allows to characterize the space in terms of statistical decomposition, and see how this differs from other decompositions.
 
@@
 The first tells us that, during mixing, the entropy cannot decrease, and it stays the same if and only we are mixing an ensemble with itself.
-The second tells us that the most the entropy can increase is given by the choice between the two ensembles.\footnote{It can be shown that if an upper bound that only depends on the mixing coefficients exists, then it must be the Shannon entropy.}
+The second tells us that the most the entropy can increase is given by the choice between the two ensembles.\footnote{The expression for the Shannon entropy is actually derived instead of imposed axiomatically.} We define two ensembles to be \textbf{orthogonal} when they maximally increase the entropy during mixing.
+This recovers the standard notion of orthogonality: two classical ensembles are orthogonal when they have disjoint support, and therefore $\int_M \rho_1 \rho_2 d\mu = $; two quantum ensembles are orthogonal when they are defined on orthogonal subspaces, and therefore $\tr(\rho_1 \rho_2) = 0$.
 
@@
 
-Another important feature is that if two ensembles saturate the upper bound, then they are orthogonal in both classical and quantum mechanics.
-Therefore we define two ensembles to be \textbf{orthogonal} if their mixture leads to a maximal entropy increase.
-An additional axiom is valid for orthogonal ensembles.
-We say an ensemble is a component of another ensemble if the second can be expressed as a mixture of the first with some other ensemble.
-On physical grounds, two orthogonal ensembles $\ens_1, \ens_2 \in \Ens$ are perfectly distinguishable and must have no common component.
-That is, there is no $\ens_3 \in \Ens$ that is a component of both $\ens_1$ and $\ens_2$.
-Note that the converse is not generally true: two pure states in quantum mechanics have no components in common, but are not necessarily orthogonal.
-We find that an ensemble space where no common component implies orthogonality can be described by a space of classical probability measures.
-Non-classicality, then, is exactly the failure of this property.
+Another crucial feature is that two ensembles maximize the entropy increase if and only if they are orthogonal: two classical ensembles maximize the entropy increase when they have disjoint support, and therefore $\int_M \rho_1 \rho_2 d\mu = 0$; two quantum ensembles when they are defined on orthogonal subspaces, and therefore $\tr(\rho_1 \rho_2) = 0$.
+This corresponds to the case when the two ensemble are made of instances that are mutually exclusive.
+Therefore we define two ensembles to be \textbf{orthogonal}, noted $\ens[a] \ortho \ens[b]$, if their mixture leads to a maximal entropy increase.
+The following additional axiom is valid for orthogonal ensembles
+\begin{enumerate}
+	\setcounter{enumi}{2}
+	\item mixtures preserve orthogonality: $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$
+\end{enumerate}
+as mixing elements distinguishable from those of ensemble $\ens[a]$, still gives elements distinguishable from ensemble $\ens[a]$.
+
+It can be shows that orthogonal ensembles are separate, but the converse is not true.
+In quantum mechanics, for example, all pure states are separate (they cannot be decomposed and therefore cannot have components in common) but not all are orthogonal.
+In classical spaces, however, all separate ensembles are also orthogonal.
+Understanding the relationships between these two properties, then, is key in understanding the difference between classical and quantum mechanics, and what can happen in other physical theories.
 
 \begin{tcolorbox}[colback=white, colframe=black]
-	Existence of entropy $\Rightarrow$ Strictly concave function whose upper bound defines orthogonality.
+	Existence of entropy $\Rightarrow$ Strictly concave function whose upper bound defines orthogonality. Statistical mixtures preserve orthogonality.
 \end{tcolorbox}
@@
 
-Each structure is important, but the balance between them is even more important.
-The key is to understand, on physical grounds, where each problem lies and therefore use the appropriate tool.
-For example, one may focus on the convex/linear structure and may be tempted to close the space under infinite convex/linear combinations.
-Mathematically, it would be very convenient.
-However, some infinite mixtures lead to physically ill-defined systems, such as a particle with an infinite average position.
-It is the connection to experimental verification, the topology, that will tell us which limits are allowed and which are not.
-Conceptual clarity on these details is essential to find the correct generalization.
-A full examination of these issues is beyond the scope of this article, and we refer to our larger, and future, work.
-However, there is one key issue that we need to discuss as, we found, it was the main conceptual obstacle to progress.
+Before introducing the actual measures, we want to note that there is a delicate balance between these structures which is important at both a conceptual level and at a more mathematical level.
+Part of the difficulty is to recognize which standard modes of understanding the problem should be kept, as the they generalize nicely, and which should be discarded.
+Let us go through some key examples.
 
-On the convex structure we can define \textbf{pure states} as the ones that cannot be decomposed into other ensembles.
+Those that focus on the convex/linear structure may be tempted to close the space under infinite convex/linear combinations.
+It would be very convenient mathematically, and, in fact, this is what happens in Hilbert spaces.
+However, this closure leads yo physically ill-defined systems, such as a particle with an infinite average position or undefined energy.(CITE Hilbert) It is the connection to experimental verification, the topology, that will tell us which limits are allowed and which are not.
+Therefore, the lack of infinite convex is a crucial feature of the space, and convex sets in ensemble spaces are to be understood as closed convex sets, so that closure on the topology includes those, and only those, infinite mixtures that are physically meaningful.
+
+If we focus on probability, we are tempted to extend the intuition from classical mechanics and it is useful to understand exactly why and where it fails.
+On a convex structure we can define the extreme points, those that cannot be further decomposed.
+Therefore it is tempting to define \textbf{pure states} as those ensembles that cannot be further decomposed into other ensembles.
 For example, the pure states for a six-sided die are the ensembles perfectly prepared respectively on the six possible outcomes.
 All ensembles are convex combinations $\sum_{i=1}^{6} p_i \ens_i = 1$ of those pure states $\ens_i$.
-Mathematically, the pure states are the extreme points of the space and all other points can be understood as convex combinations over them.
-This is exactly how classical discrete spaces work and there is a temptation to think this is how everything works.
-For example, in classical mechanics the pure states would be perfect preparations of position and momentum, and ensembles are probability distributions.
-Not only is this problematic mathematically, it is conceptually not sustainable.
+This is exactly how classical discrete spaces work and much of our intuition for probability comes from there.
 
-On physical grounds, we cannot consistently prepare classical particles perfectly, therefore there is no ensemble corresponding to a perfect preparation of a point of phase space.
-The mathematical facts that delta Dirac are not proper functions, their support has an empty interior, have minus infinite entropy, and so on, are simply a consequence of our conceptual problem.
-One may concede that the ensembles do not exist physically, but may still want them to consider them as the limit of infinitesimal decomposition: classical ensembles can be decomposed into more and more precise ones, and the points are the limit of this decomposition.
-Probability distributions are defined on these limits.
-This kind of works in classical mechanics, but in general.
+This approach, however, already becomes problematic in classical mechanics.
+Since we can take uniform distributions over any region of phase space, we can imagine to take a sequence of more and more refined distributions around a value of position and momentum.
+Therefore we may be tempted to say that the points of phase space represent perfectly prepared systems, and every other probability is a distribution on these space.
+This is physically unsound.
+First of all, we cannot consistently prepare position and momentum with infinite precision, therefore those ensembles do not exist even exist as an idealization.
+Secondly, the entropy of those ensembles, if they existed, would be minus infinity, which is a problem for any thermodynamic treatment.
+To make this intuition consistent with the math would require each point in phase space to be topologically isolated, which would extend the $\sigma$-algebra to the power set of $\mathbb{R}^n$, which means there would be no consistent way to define a volume.
+The fact that the math breaks down is simply a symptom that our physical problem is ill-specified.
 
-In quantum mechanics we do have pure states, and these are ensembles that cannot be further decomposed into finer ensembles.
-Therefore an eigenstate of position is not a limit of decomposition, a limit in the convex structure.
-It is the limit where we are preparing a pure state on a smaller and smaller region of space.
-It is the limit of a sequence of pure states, a limit in the topology.
-Physically, more than mathematically, it is a completely different thing.
-It is only when we realized this difference that a clean route to generalization presented itself.\footnote{The fact that pure states can be seen, in some sense, as superposition of eigenstates of position, makes the matter even more confusing.}
+In quantum mechanics, things are even worse.
+Pure states are indeed extreme points of the convex set.
+However, we also have discrete spectra.
+Here the temptation is, like we ``did'' in classical mechanics, to consider wave-functions all concentrated at a single value of position.
+While one may think this is the analogue of classical mechanics, it is not.
+We can, in fact, take a wave-function that is uniformly distributed on every finite range of position, and we can imagine to shrink that range.
+However, this is a sequence of pure states.
+All pure states have zero entropy, and the reduced spread on position will mean an increased spread on momentum.
+That is, we are not providing a sequence of ensembles at ever greater prevision, like in the classical case.
+In other words, the eigenstates of position in quantum mechanics, even if we wanted to say they were actual states, they would not be the limit of infinitesimal decomposition (i.e.
+a limit in the convex structure) but rather the limit of a path over pure states (i.e.
+a topological limit).
 
 The insight is that probability distributions for a quantity are not defined by statistical mixtures of perfect preparations (i.e.
-a convex combination of extreme points), but by assigning a probability over each experimentally verifiable (i.e.
-open) interval of the possible values of said quantity (its spectra).
+convex combinations of extreme points), but by assigning a probability over each experimentally verifiable (i.e.
+open) interval of the possible values of said quantity (its spectrum).
 The two coincide in classical discrete spaces, which are the least physically interesting case as they do not even allow classical particle mechanics.
 Equivalently, \textbf{mixtures are about preparations, not measurements, and probability distributions are about measurements, not preparations}.
-Since they can only coincide in classical spaces, a generalized notion of probability for physical theories must resolve and characterize the distinction.
+In quantum mechanics, an equal mixture of spin up and spin left, for example, does not correspond to a 50\% probability of measuring spin up: the mixing coefficients and the outcome probability are different.
 
@@
 The convex structure captures statistical mixing and is responsible for all linear structures of physical theories.
-The entropic structure captures the variability within an ensemble and is ultimately responsible for all geometric structure.
-We now concentrate on recovering the measure theoretic structures associated to probability and state counting.
+The entropic structure captures the variability within an ensemble and is ultimately responsible for all geometric structures.
+Since the physical problem is indivisible, there is an interplay across these structure which goes beyond this article and we refer to our larger work.
 
@@
 
-Note that the mixing coefficient $p$ is connected to probability but it is not itself probability in the standard sense.
-It represents a parameter for the mixing states, mixing preparations, while probability is typically understood as the chance of something happening, the likelihood of a final event.
-Mathematically, the probability measure is a function from an event, which can be understood as a collection of fully distinct outcomes, to a number from zero to one.
-If we are mixing ensembles that represent fully distinct outcomes (i.e.
-they are orthogonal), then the mixing $p$ is a probability in the standard sense.
-If we are mixing ensembles that do not represent fully distinct outcomes, that are not orthogonal like spin up and spin left, the probability of getting one of the outcomes is not the mixing coefficient $p$.
-
-We essentially have two ways in which two events can be compared.
-We have \textbf{disjoint events}, those that represent disjoint sets of different states, like spin up and spin down; we also have \textbf{orthogonal events}, those that represent represent orthogonal sets of states, like spin up and spin down.
-The standard notion of probability applies only when considering orthogonal events.
-Classical probability rests on the assumption that all disjoint events are orthogonal.
-That is, that all states can be perfectly distinguished.
-Quantum mechanics breaks this assumption, and therefore we need a probability theory that can deal with disjoint but non-orthogonal events.
-
-Let us concentrate on the following question: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
+We have clarified that mixing coefficients are not the same as probabilities of outcomes.
+Unlike most of the literature on probability, we will concentrate on mixtures and, particularly, on the following question: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
 How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
@@
 
+In the classical case, the fraction capacity reduces to a probability measure.
 Note that a probability satisfies all the properties of a fraction capacity, but it also satisfy on additional property: additivity over disjoint sets.
-That is, $\mu(A \cup B) = \mu(A) + \mu(B)$ if $A \cup B = \emptyset$.
+That is, if $A \cup B = \emptyset$ then $\mu(A \cup B) = \mu(A) + \mu(B)$.
+Adding finite additivity over disjoint sets is enough as continuity will extended the property to additivity over countably many disjoint sets.
+Conversely, having countable additivity over disjoint sets implies sub-additivity and continuity from below and above.
+
+While additivity over disjoint set holds in classical probability, it does not hold in quantum mechanics
+
  This is not true in general, as we can see in a simple example from quantum mechanics.

### 2025-05-01 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: TOC skeleton

@@
 \maketitle
+
+\section{TOC}
+
+Overall goal: to create a generalization of probability theory and measure theory for general physical systems.
+
+General question: how do we represent physical states? Different types of representations: talk about examples, ....
+
+-- General problem
+
+For this work: assume we have a locally convex topological vector space V. E is a bounded convex subset of V. Equipped with an entropy. Topology generated by entropic balls?
+
+Separateness and orthogonality
+
+Statistical quantities
+
+Show that they reduce to standard things in classical mechanics and quantum mechanics.
+
+-- Connection to Choquet theory
+
+Choquet theory - in the case that the closure of the space is compact -> representation through measures. May not be unique
+
+Choquet simplex. Uniqueness of decomposition and separateness/orthogonality.
+
+-- Connection to signed probability
+
+TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
+
+Characterize signed measures as an affine representation of the ensemble space (i.e. bounded convex subset of a locally convex topological vector space).
+
+Show that the probability is signed if and only if multiple decomposition, non-additive, no Choquet simplex
+
+Signed probability is unique, but does not characterize the convex set (ensemble space).
+Note: the other measures are defined on the extreme points.
+Knowing the extreme points means you know the convex space (i.e.
+take the hull).
+The non-additive measure does not characterize the extreme points, so it doesn't really characterize the space.
+
+-- Our non-additive measure approach
+
+State capacity
+
+Fraction capacity. Show uniqueness. Also show uniqueness on extreme points.
+
+Fraction capacity as supremum of all possible measures
+
+Additivity of fraction capacity (and state capacity).
+
+Single decomposition and classical spaces and measurement contexts
+
+Show that they reduce to standard things in classical mechanics and quantum mechanics.
+
+-- Connection to fuzzy logic?
+
+Literature search.
+
+Possible connection to
+* Choquet integral
+* Sugeno integral
+
+? Space of fuzzy measure like we have a space of additive measures?
 
@@
 
-The insight is that probability distributions for a quantity are not defined by statistical mixtures of perfect preparations (i.e.
-convex combinations of extreme points), but by assigning a probability over each experimentally verifiable (i.e.
-open) interval of the possible values of said quantity (its spectrum).
-The two coincide in classical discrete spaces, which are the least physically interesting case as they do not even allow classical particle mechanics.
-Equivalently, \textbf{mixtures are about preparations, not measurements, and probability distributions are about measurements, not preparations}.
-In quantum mechanics, an equal mixture of spin up and spin left, for example, does not correspond to a 50\% probability of measuring spin up: the mixing coefficients and the outcome probability are different.
+The key insight is that \textbf{mixing coefficients and probabilities are different in general} and they coincides only when a particular ensemble can be understood as a statistical mixtures of perfect preparations of a given quantity (i.e.
+of orthogonal ensembles).
+This always happens in the classical case and over measurement contexts in the quantum case, but is not the general case.
+For example, an equal mixture of spin up and spin left, for example, does not correspond to a 50\% probability of measuring spin up.
+The key to develop a general theory of probability, then, is to characterize the mixing coefficients first and understand when they recover probabilities of outcomes.
 
@@
 
-We have clarified that mixing coefficients are not the same as probabilities of outcomes.
-Unlike most of the literature on probability, we will concentrate on mixtures and, particularly, on the following question: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
+Given that mixing coefficients have a more general applicability than probability of outcomes, let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
 How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
@@
 
-One can then that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
+One can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
 \begin{enumerate}
-	\item non-negative and unit bounded - $0 \leq \frcap_{\ens}(A) \leq 1$
+	\item non-negative and unit bounded - $\frcap_{\ens}(A) \in [0,1]$
 	\item monotone - $A \subseteq B \implies \frcap_{\ens}(A) \leq \frcap_{\ens}(B)$
@@
 The limit of these sequences must agree with the fraction capacity of the limit.
+
+If not for the lack of additivity over disjoint sets, the fraction capacity would be a probability measure.
+However, the difference between fraction capacity and probability goes beyond that.
+While the probability measure is defined over sets of outcomes, the fraction capacity is defined over ensembles, over probability distributions.
+Conceptually the two are different object, but they, as we will see later, linked.
+Suppose $\Ens$ is the classical ensemble space given by the set of probability distributions over a sample space $\Omega$.
+Let $U \subseteq \Omega$ be an event, let $A_U \subset \Ens$ be the set of all probability distributions with support within $U$ and $\ens \in \Ens$ a probability distribution.
+Then we can show that $\frcap_{\ens}() $
 

### 2025-05-02 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reworked the introduction

@@
 \section{Introduction}
+
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories.
+One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
+
+We reformulate the problem in the following way: given the space of all possible statistical ensembles allowed by a statistical theory, what measure theoretic tools can we use to represent each ensemble?
+If they are not always available, in what circumstances and with what limitations are they available?
+Is a suitable generalization possible, and what mathematical tools would need to be developed?
+
+In this paper, we will define a space of ensembles as a convex bounded set of a Hausdorff locally convex topological vector space, upon which an entropy function is defined.
+These mathematical properties are physically grounded in the ability to experimentally define ensemble (leading to a topological space), to define statistical mixtures (leading to a convex space), to quantify the variability of the elements of the ensembles (leading to an entropy functions) and to having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
+The question is, then, what type of measure theoretic tools can we use to quantify the number of states and to represent each ensemble in the space.
+Both are needed to be able to define a physically meaningful notion of probability density.
+
+We will first review some results from Choquet theory, which is, to our knowledge, the most general approach to represent elements of a convex set using additive probability measures.
+The key insight is that, at least in the compact case, such representation is always possible, but it is not unique.
+Uniqueness requires the space to be a Choquet simplex, which is a generalization of the standard finite-dimensional simplex.
+Quantum theory does not fall in this case: it is precisely the multiple decomposition of mixed states into pure states that prevents a quantum ensemble space to be simplex.
+
+We will also review some results from the use of signed probability measures.
+While the most known example is the use of Wigner functions, there are actually several ones and there is also work on developing a general theory of signed probability.
+Given that the space of signed probability measure is, unlike a simplex, a vector space, we can find an affine map between signed probability measures and the embedding vector space.
+This gives us a unique representation of each ensemble in term of a singed probability measure.
+However, unlike in the Choquet case, the measure is not supported by the set of extreme points of the space.
+That is, there is no unique or privileged choice for this representation.
+Moreover, since the map is at the level of the full vector spaces, there is no general way to characterize which subset of signed probability measures correspond to ensembles.
+
+At this point, we will show that we can define two types of measures: the state capacity, which generalizes the notion of count of states, and a fraction capacity for each ensemble, which generalizes the notion of probability.
+Both are non-negative sub-additive monotonic set functions, which the second being unit bounded.
+These are defined on the ensemble space itself, which means the domain is uniquely defined by the ensemble space.
+Moreover, in the compact case, these measure can be restricted on the extreme points and still retain uniqueness.
+The fraction capacity, in fact, reduce to the supremum of all Choquet measures over the extreme points.
+The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to the unique measure, and is therefore additive.
+The state capacity, instead, recovers the Liouville measure in the classical case and the dimensionality of the Hilbert space in the quantum case.
+
+Having shown that a general approach based on non-additive measures is possible, we lay down a series of questions to explore in the future.
+The biggest one is whether it is possible to generalize notions such as integration, derivation and expectation to this non-additive case.
+It should be noted that non-additive measures are used in the context of decision theory, so there may be a connection between the use of non-additive measures in quantum theory and interdependent complimentary choices.[TODO: find citation that sub-additive is when choices are complementary] Both the Choquet and the Sugeno integral, for example, are defined over a monotonic set-functions, so there is a chance that these tools may turn out to be useful.
+We see this work as starting this line of research, which will take more effort to investigate.
+
+\section{Old introduction}
 

### 2025-05-02 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: adding ensemble space section

@@
 We see this work as starting this line of research, which will take more effort to investigate.
+
+\section{Ensemble space}
+
+Conceptually, a statistical ensemble is a collection of outputs, taken at once, of a reproducible preparation procedure for a physical system.
+In the classical case, this corresponds to a probability distribution over a discrete set of outcomes (i.e.
+a fair coin) or over phase space (i.e.
+the symplectic manifold represent position and momentum for a classical particle).
+In the quantum case, this corresponds to a mixed state (i.e.
+a positive semi-definite self-adjoint operator of trace one).
+An ensemble space is the collection of all possible ensembles defined in a physical theory, together with a basic mathematical structure that all physical theories must possess.
+The development of a full theory of ensembles goes beyond the scope of this work, and we refer to the appendix and our ongoing work (cite).
+
+For the purpose of this work, we will assume an ensemble space $\Ens$ is a convex subset of a Hausdorff locally convex topological vector space $V$.
+The topology is the physically natural topology, where each open set represents a verifiable statement.
+The convex structure represents statistical mixing.
+For example, $p \ens[a] + \bar{p} \ens[b]$ represent the statistical mixture of two ensembles $\ens[a]$ and $\ens[b]$ with probability $p$ and $\bar{p} = 1 - p$.
+Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i}^{\infty} p_i \ens_i$ converges in the space or not.
+We also assume that there is a countable family of affine continuous functions $F_i : \Ens \to \mathbb{R}$ that represent statistical quantities and fully identify each ensemble.
+That is, $F_i(p \ens[a] + \bar{p} \ens[b]) = p F_i(\ens[a]) + \bar{p} F_i(\ens[b])$ and $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.
+For example, we can imagine $F_i$ representing the expectation of all polynomials of position and momentum.\footnote{In this case, the possible probability distributions would be restricted to those that go to zero at infinity on all variables faster than any polynomial.
+We will leave these technical details out of the main section of the text as to not overwhelm the reader.} Mathematically, these statistical quantities can be used to define semi-norms on the space, giving us a locally convex topology.
+
+Additional, each ensemble space $\Ens$ is equipped with a continuous entropy function $S : \Ens \to \mathbb{R}$. The entropy of a statistical mixture satisfies the following bounds
+\begin{equation}
+	\begin{aligned}
+		p S(\ens[a]) + (1-p) &S(\ens[b]) \leq S(p \ens[a] + (1-p) \ens[b]) \\
+		&\leq I(p,\bar{p}) + p S(\ens[a]) + (1-p) S(\ens[b])
+	\end{aligned}
+\end{equation}
+where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.
+The bounds come from assuming that the entropy quantifies the variability of each ensemble.
+Intuitively, the variability cannot decrease during mixing, and will have an upper bound when mixing two ensemble that are non-overlapping, or orthogonal.
+In fact, we define two ensembles $\ens[a] \ortho \ens[b]$ to be orthogonal when they mixture gives a maximal entropy increase.
+This does recover the notion of orthogonality in the corresponding classical and quantum vector spaces of statistical ensembles.
+Lastly, mixtures preserve orthogonality in the sense that mixing non-overlapping ensembles still gives non-overlapping ensembles.
+More precisely, $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$.
+
+To summarize, the above mathematical structure represents the requirements that ensembles must be experimentally well-defined, allow statistical mixture, can be identified with statistical quantities and provide a well-defined entropy.\footnote{While the full physical justification of the above mathematical structure is beyond the scope of this work, it is important to us that such justification exists and this physically straight forward.
+This is the essence of what we call Physical Mathematics.
+We will summarize some of these justifications in the appendix, leaving the full and more updated discussion in our work in progress.} We now turn to the question of how we can represent both the entropy and each ensemble in terms of some measure theoretic structure.
 

### 2025-05-12 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: more work on the rewrite

--- 2025-05-01 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-Our current understanding of probability is given in terms of the measure theoretic formalization due to Kolmogorov, and its various philosophical interpretations.
-However, probability in quantum mechanics, as it is widely known, is not Kolmogorovian, in the sense that it provides results that are not reproducible in standard probability spaces.
-It follows that our mathematical axioms for probability and their interpretations are insufficient for physical theories.
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories.
+One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
 
-An alternative measure theoretic strategy, that has been successful is quantum mechanics, is the use of quasi-probability distributions: measures that are still unitary over the whole space, but can have negative values in small regions.
-However, these haven't resulted in an alternative axiomatic approach with suitable interpretations for at least, in our mind, three reasons.
-First, it is not clear how negative probability should be interpreted.
-Second, it is not clear how to characterize the space of valid quasi-probability distributions (e.g.
-there is no simple rule to specify which $\rho(q,p)$ are valid Wigner functions).
-Third, there are multiple options for quasi-probability distributions (e.g.
-Wigner function, Husimi distribution, Sudarshan–Glauber distribution).
-This lack of foundational clarity presents an additional problem: how do we know whether future physical theories will not require yet another extension?
+We reformulate the problem in the following way: given the space of all possible statistical ensembles allowed by a statistical theory, what measure theoretic tools can we use to represent each ensemble and the geometric properties of the overall space?
+Why exactly classical probability measures fail to capture completely quantum states?
+What suitable generalizations are possible, and what mathematical tools would need to be developed?
+
+In this paper, we will characterize a generic space of ensembles as a convex bounded set of a Hausdorff locally convex topological vector space, upon which an entropy function is defined.
+These mathematical properties are physically grounded in the ability to experimentally define ensemble (leading to a topological space), to define statistical mixtures (leading to a convex space), to quantify the variability of the elements of the ensembles (leading to an entropy function) and to having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
+The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the number of orthogonal subspaces of an operator)?
+What measure theoretic tool can represent each ensemble?
+If we are to define a physically meaningful notion of probability density, both of these questions must be answered.
+
+We will first review some results from Choquet theory, which is, to our knowledge, the most general approach to represent elements of a convex set using additive probability measures.
+The key insight is that, at least in the compact case, such representation is always possible, but it is not unique.
+Uniqueness requires the space to be a Choquet simplex, which is a generalization of the standard finite-dimensional simplex and is a particularly shaped subset of a vector space.
+The space of quantum ensembles is not a simplex for one key reason: it allows multiple decompositions in terms of pure states.
+These insights will allow to better understand exactly the limits of classical probability to describe ensembles in physical theories.
+
+We will also review some results from the use of quasi-probability measures.
+While the Wigner function represents the most known instance, there are different choices.
+Note that the space of signed probability measures over a set $K$ is a vector space, so one can find an affine map between a suitable space of signed probability measures and the embedding vector space.
+Each ensemble, then, is uniquely represented with a signed probability measure.
+However, unlike in the Choquet case, the measure is not supported by the set of extreme points of the ensemble space.
+That is, there is no unique or privileged choice for this representation.
+Moreover, since the map is at the level of the full vector space, not just the subset with valid ensembles, there is no general way to characterize which subset of signed probability measures correspond to ensembles.
+Therefore, while signed measures are a powerful tool for calculation in specific problems, they cannot give us a general principled representation of the physical theory.
+
+The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
+From the entropy, we can define the state capacity, which generalizes the notion of count of states.
+For each ensemble, we can define the fraction capacity, which generalizes the notion of probability.
+Both are non-negative sub-additive monotonic set functions, with the second being unit bounded.
+These measures are always unique and, since they are defined on the ensemble space itself, always have provide a physically meaningful representation.
+Moreover, in the compact case, these measure can be restricted on the extreme points while retaining uniqueness.
+(TODO: prove)
+
+The fraction capacity is the supremum of all possible measures for the same ensemble.
+The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure, and is therefore additive.
+The state capacity, instead, corresponds to the minimum of the entropy of all possible representations.
+It recovers the Liouville measure in the classical case and the dimensionality of the Hilbert space in the quantum case.
+
+Having shown that a general approach based on non-additive measures is possible, we lay down a series of questions to explore in the future.
+The biggest one is whether it is possible to generalize notions such as integration, derivation and expectation to this non-additive case.
+It should be noted that non-additive measures are used in the context of decision theory, so there may be a connection between the use of non-additive measures in quantum theory and interdependent complimentary choices.[TODO: find citation that sub-additive is when choices are complementary] Both the Choquet and the Sugeno integral, for example, are defined over a monotonic set-functions, so there is a chance that these tools may turn out to be useful.
+We see this work as starting this line of research, which will take more effort to investigate.
+
+A final remark before we start.
+Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects.
+Therefore, it is important to us that this work is a piece of a much larger puzzle, which connects cleanly and directly to  other areas, such as information geometry, functional analysis, quantum logic and so on.
+It is the hallmark of a true fundamental structure for physics that such connections can be easily made.
+Unfortunately, the nature of scientific publication requires us to publish a narrow slice of the work to fit the scope of a particular journal.
+Ironically, this means that the most important motivations for our approach are out of scope of this article, as we will concentrate only to the connections to measure theory.
+
+\section{Ensemble space}
+
+Conceptually, a statistical ensemble is a collection of outputs, taken at once, of a reproducible preparation procedure for a physical system.
+In the classical case, this corresponds to a probability distribution over a discrete set of outcomes (i.e.
+a fair coin) or over phase space (i.e.
+the symplectic manifold representing position and momentum for a classical particle).
+In the quantum case, this corresponds to a mixed state (i.e.
+a positive semi-definite self-adjoint operator of trace one).
+An ensemble space is the collection of all possible ensembles defined in a physical theory, together with a basic mathematical structure that all physical theories must possess.
+The development of a full theory of ensembles goes beyond the scope of this work, and we refer to the appendix and our ongoing work (cite).
+
+For the purpose of this paper, we will assume an ensemble space $\Ens$ is a convex subset of a Hausdorff locally convex topological vector space $V$.
+The topology is the physically natural topology, where each open set represents a verifiable statement.
+The convex structure represents statistical mixing.
+For example, $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture of two ensembles $\ens[a]$ and $\ens[b]$ with probability $p$ and $\bar{p} = 1 - p$.
+Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i}^{\infty} p_i \ens_i$ converges in the space or not.
+We also assume that there is a countable family of affine continuous functions $F_i : \Ens \to \mathbb{R}$ that represent statistical quantities and fully identify each ensemble.
+That is, $F_i(p \ens[a] + \bar{p} \ens[b]) = p F_i(\ens[a]) + \bar{p} F_i(\ens[b])$ and $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.\footnote{It is an open question whether all ensemble spaces must possess this property.} For example, we can imagine $F_i$ representing the expectation of all polynomials of position and momentum.\footnote{This requires all probability distributions to converge to zero at infinity faster than any polynomial.} Mathematically, these statistical quantities can be used to define semi-norms on the space, giving us a locally convex topology.
+
+Additional, each ensemble space $\Ens$ is equipped with a continuous entropy function $S : \Ens \to \mathbb{R}$. The entropy of a statistical mixture satisfies the following bounds
+\begin{equation}
+	\begin{aligned}
+		p S(\ens[a]) + (1-p) &S(\ens[b]) \leq S(p \ens[a] + (1-p) \ens[b]) \\
+		&\leq I(p,\bar{p}) + p S(\ens[a]) + (1-p) S(\ens[b])
+	\end{aligned}
+\end{equation}
+where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{In the larger work, we can show that the Shannon entropy is the only upper bound possible.} The bounds come from assuming that the entropy quantifies the variability of each ensemble.
+Intuitively, the variability cannot decrease during mixing, and will have an upper bound when mixing two ensembles that are non-overlapping, or orthogonal.
+In fact, we define two ensembles $\ens[a] \ortho \ens[b]$ to be orthogonal when they mixture gives a maximal entropy increase.
+This does recover the notion of orthogonality in the corresponding classical and quantum vector spaces of statistical ensembles.
+Lastly, mixtures preserve orthogonality in the sense that mixing non-overlapping ensembles still gives non-overlapping ensembles.
+More precisely, $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$.
+
+To summarize, the above mathematical structure represents the requirements that ensembles must be experimentally well-defined, allow statistical mixture, can be identified with statistical quantities and provide a well-defined entropy.\footnote{While the full physical justification of the above mathematical structure is beyond the scope of this work, it is important to us that such justification exists and this physically straight forward.
+This is the essence of what we call Physical Mathematics.
+We will summarize some of these justifications in the appendix, leaving the full and more updated discussion in our work in progress.} We now turn to the question of how we can represent both the entropy and each ensemble in terms of some measure theoretic structure.
+
+\section{Additive measure and Choquet theory}
+
+Since ensemble spaces are a convex subset of a locally convex topological vector space, they can be connected to Choquet theory.
+First of all, let $E$ be a convex subset of a locally convex space $V$.
+We say that $\ens \in E$ is represented by a probability measure $\mu : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_X f d\mu$ for every continuous linear functional $f$ on $V$.
+The Choquet theorem states:
+\begin{thrm}[Choquet]
+	Let $E$ be a metrizable compact convex subset of a locally convex space $V$, and let $\ens \in E$.
+Then there exists a probability measure $\mu_{\ens}$ on $X$ which represents $\ens$ and is supported by the extreme points of $X$.
+\end{thrm}
+
+First, let us understand the physical significance.
+An ensemble $\ens$ can be seen as a probability measure $\mu$ over the ``pure states'' (i.e.
+the extreme points of $E$) if the value for every statistical quantity $f$ for $\ens$ can be recovered as an expectation over the pure states using the probability measure.
+That is, all statistical properties of $\ens$ can be recovered by the probability measure.
+This is what allows to think go $\ens$ as a probability distribution.
+Note, however, that the Choquet theorem does not guarantee uniqueness.
+The probability measure is unique if and only if $E$ is a Choquet simplex, which is a generalization of the finite dimensional simplex.
+The characterization of the Choquet simplex is fairly technical, so we will simply define the Choque simplex by the property of having a single decomposition in terms of extreme points.
+
+We should now look at the requirements of the theorem, to see whether we can always apply it to an ensemble space.
+First of all, and ensemble space $\Ens$ is a convex subset of a convex sapce $V$.
+Moreover, since every Hausdorff first-countable topological vector space is metrizable, $\Ens$ is metrizable.
+However, $\Ens$ is not necessarily compact.
+In fact, $\Ens$ is not necessarily a closed subset of $V$.
+Consider the space of probability measures over $\mathbb{R}$.
+Since $\mathbb{R}$ is not compact, the space of probability measure over $\mathbb{R}$ cannot be compact.\footnote{TODO: Check with mathematicians} Moreover, the set of extreme points is given by the Dirac measures (i.e.
+probability one over a single point), which would have minus infinite entropy, and therefore cannot be part of the ensemble space.
+Therefore the theorem does not apply in general.
+
+We can apply Choquet theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex} and metrizable.
+In this case, we are guaranteed that all ensembles allows a representation through a probability on points that may corresponds idealized ensembles (i.e.
+the Dirac measure) that are not, even conceptually, realizable.
+Many interesting cases fall into this restriction: the ensemble spaces of classical probability distribution over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
+
+Note that ensembles over classical spaces will be represented by a unique probability measure, therefore the corresponding closure of the ensemble space is a Choquet simplex.
+However, quantum mixed states can be decomposed in multiple ways in terms of pure states, and therefore the closure of a quantum ensemble space is not a simplex.
+
+% space is metrizable https://proofwiki.org/wiki/Birkhoff-Kakutani_Theorem/Topological_Vector_Space
+
+Chouqet theory tells, in the realm where it can be applied, that classical probability fails in quantum mechanics precisely because the ensemble space of a quantum system allows multiple decomposition.
+
+\section{Quasiprobability}
+
+Another way to represent quantum ensembles is through the use of quasiprobability distributions.
+Given a space $X$, a measure $\mu$ is affine if $\mu(X)$.
+All probability measures are affine measures, but the reverse is not true as affine measures can be negative on particular regions.
+The Wigner function, for example, can be negative over small regions of phase space.
+
+Given a space $X$, the space of all affine measure is a vector space, not just a convex subset as in the case of the space of probability measures.
+This means that, given an ensemble space, we may find a space $X$ such that the space of affine measures over $X$ is isomorphic to the embedding vector space $V$ as topological vector spaces.
+In this case, any element $v\in V$ can be represented uniquely by an affine measure $\mu_{v}$, regardless of whether they are actual ensembles or not.\footnote{The question of whether such an $X$ always exists is left open.
+We suspect a link exists between the statistical quantities $F_i$ that fully identify each ensemble and the space $X$.} Because the statistical quantities $F_i$ are affine functions, we will have $F_i(v) = \int_X F_i d\mu_{v}$.
+
+In general, since the ensemble space is bounded along every direction, not all affine measure will correspond to an ensemble space.
+Whether all probability measure correspond to a valid ensemble, or whether affine measures are even used, correspond to relationships between $\Ens$ and the space of probability measures $M_p(X)$.
+If $\Ens \subset M_p(X)$, then all states are represented by a probability measure.
+However, this is not a ``probability'' in the standard sense of the word.
+For example, the Husimi Q representation is always non-negative at every point, but two points to do not represent mutually exclusive events as they are not orthogonal (i.e.
+their inner product is non-zero and they do not maximize the entropy increase during mixture).
+Mathematically, the issue is that the points $X$ of the Husimi Q representation are not the extreme points of the closure ensembles space $\overline{\Ens}$.
+In general, $\Ens$ and $M_p(X)$ will overlap, but neither will contain the other.
+For example, in the Wigner representation the only pure states (i.e.
+the extreme points of $\Ens$) represented by a probability measure are the Gaussian wave-packets.
+This means that any other non-negative Wigner function will be in the convex hull of Gaussian wave-packets, and therefore will allow a decomposition in terms of them.
+
+Note that if $\overline{\Ens} = M_p(X)$, then the space of the extreme point is exactly the space of Dirac measures, and therefore $X$ corresponds to the pure states.
+Alternatively, either one ensemble is represented by an affine measure, or not all probability measures can be understood as mixtures of extreme points.
+If $\overline{\Ens}$ is compact, then this is exactly when $\overline{\Ens}$ is not a simplex and multiple decompositions are allowed.
+
+This alternative way of looking at measure theoretic representations, then, is consistent with the previous one.
+While quasiprobability distributions are very useful in practice, we want to stress that there is no single choice for $X$.
+To solve a specific problem, this may actually be useful as one can choose the most convenient $X$ for the problem.
+To develop a general theory for physics, however, this is not useful as each representation is not general enough.
+
+\section{Old introduction}
 
@@
 This points to a possible generalization of measure theoretic calculus to a sub-additive case that would be guaranteed to work on all physical theories.
-
-\section{Quick review of probability theory}
-
-The standard way to characterize probability is through measure theory.
-Given a sample space $\Omega$, which represents the possible outcomes, and a $\sigma$-algebra $\Sigma_{\Omega}$, which represents the events, we define a probability measure $p : \Sigma_{\Omega} \to [0,1]$.
-Note that the probability is assigned to events, not the outcomes.
-This is what allows to treat discrete and continuous spaces within the same framework.
-In physics, all sample spaces are ultimately topological spaces, and the $\sigma$-algebra is typically the Borel algebra (i.e.
-the smallest $\sigma$-algebra that contains all open sets).
-
-In classical mechanics, the sample space is phase space $M$ given by all possible values of position and momentum for all particles under considerations.
-Phase space is equipped with another measure, the Liouville measure $\mu : \Sigma_M \to [0, +\infty]$, that returns volume of each region.
-In statistical mechanics, this volume represents the count of states and is another key ingredient.
-The probability density over states is given by the Radon-Nikodym derivative $\rho = \frac{dp}{d\mu}$ and the entropy is given by $S(\rho) = - \int_M \rho \log \rho d\mu$.
-In case of a uniform distribution $\rho_U$ with support $U$, the entropy is given by $S(\rho_U) = \log \mu(U)$.\footnote{Note that entropy, the count of states and the symplectic form that characterizes phase space are deeply intertwined, yet they require different mathematical framework to express them.
-We suspect that they are "shadows" of a more fundamental mathematical structure, which our work on ensembles space is trying to identify.}
-
-The above definitions do not apply to quantum mechanics.
-In the Hilbert space formulation, mixed states are represented by density operator (i.e.
-positive semi-definite self-adjoint operators with trace one).
-Note that if $\rho(x)$ is an integrable function such that $\int_X \rho(x) dx = 1$, the map $\psi(x) \mapsto \rho(x) \psi(x)$ is a density operator.
-Therefore the space of probability densities over position are mixed states in quantum mechanics.
-However, no joint distribution between position and momentum can be given, therefore quantum states cannot be described by the space of probability measures over a single sample space.
-
-Alternative representations over phase space do exist, and they are widely used.(CITE) The use quasi-probability measure, meaning that the measure over the full space is still unitary (i.e.
-$\mu(M) =1$).
-However, the measure can be negative for smaller regions, which means it can also be grater than $1$, and there is no simple characterization of which quasi-probability measures are allowed and which aren't.
-Moreover, there is no simple link between then entropy of the distribution and the Liouville measure.
 

### 2025-05-13 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewing section on fraction capacity

@@
 
-We can apply Choquet theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex} and metrizable.
+We can apply Choquet theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
+TODO: finalize the proof} and metrizable.
 In this case, we are guaranteed that all ensembles allows a representation through a probability on points that may corresponds idealized ensembles (i.e.
@@
 To develop a general theory for physics, however, this is not useful as each representation is not general enough.
+
+\section{Fraction capacity}
+
+TODO: intro
+
+Given that mixing coefficients have a more general applicability than probability of outcomes, let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
+How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
+For example, let $\Ens$ be the space of all probability distributions for a six-sided die.
+Let $\ens_{123456}$ be the uniform distribution over all outcomes.
+Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
+Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$, where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/6$ of $\ens$ can be constructed from $A$, but no more.
+Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
+Note that, given the uniform distribution, $1/6$ is exactly the probably for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
+This gives the basic insight for our definitions.
+
+Given a target ensemble $\ens \in \Ens$ and an arbitrary ensemble $\ens[a] \in \Ens$, we define the \textbf{fraction} of $\ens[a]$ in $\ens$ to be
+\begin{equation}
+	\fraction_{\ens}(\ens[a]) = \sup(\{ p \in [0,1] \, | \, \exists \, \ens_1 \in \Ens \text{ s.t. }  \ens = p \ens[a] + \bar{p} \ens_1 \}).
+\end{equation}
+The fraction is always well-defined because we can always write $\ens = 0 \ens[a] + 1 \ens$, therefore it must be zero or greater.
+This quantity tells us how much of ensemble $\ens$ can be constructed from $\ens[a]$.
+
+We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a Borel set of ensembles $A \subset \Ens$.
+Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.
+mixtures),\footnote{Since we work with convex sets within a topology, the hull is both the closed under the topology and convex combinations.
+That is, $\hull(A)$ is the smallest closed and convex set that contains $A$} we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
+\begin{equation}
+	\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\}).
+\end{equation}
+This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$.\footnote{The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
+fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.
+TODO \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} }
+
+One can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
+\begin{enumerate}
+	\item non-negative and unit bounded - $\frcap_{\ens}(A) \in [0,1]$
+	\item monotone - $A \subseteq B \implies \frcap_{\ens}(A) \leq \frcap_{\ens}(B)$
+	\item sub-additive - $\frcap_{\ens}(A \cup B) \leq \frcap_{\ens}(A) + \frcap_{\ens}(B)$
+	\item continuous from below and above - $\frcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \frcap_{\ens}(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
+\end{enumerate}
+
+The first property readily comes from the domain of the fraction.
+For the second, note that $\hull$ and $\sup$ are both monotone.
+For sub-additivity, note that the fraction of two components can at most sum during mixing.
+For the last, increasing and decreasing sequences strictly add or remove possible mixtures, therefore they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
+The limit of these sequences must agree with the fraction capacity of the limit.
+If not for the lack of additivity over disjoint sets, the fraction capacity would be a probability measure.
+
+Our setup is very similar to what happens in Choquet theory.
+The fraction capacity, in fact, is not just defined on the extreme points, but it is defined on the whole space of ensembles.
+In the classical discrete case, for example, if $A \subset \Omega \subset \Ens$ is an event, then $\frcap_{\ens}(A)$ will return the probability of event $A$; if $P \subset \Ens$ represents all the Poisson distributions with integer coefficient, $frcap_{\ens}(P)$ will return how much of $\ens$ can be expressed is a mixture of those distributions.
+Similarly, the probability measures in Choquet theory are, in general, defined on the whole ensemble space, and then one finds those particular ones that have support on the extreme points.
+
+We can establish a direct connection between the fraction capacity and Choquet theory, because the ability to express an ensemble $\ens$ as the convex combination of $A$ is precisely the ability to find a measure $\mu$ that has support within $A$.
+That is, $\ens = \sum_{i=1}^{\infty} p_i \ens[a]_i$ if an only if the probability measure defined by $\mu(\{\ens[a]_i\}) = p_i$ represents $\ens$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{\mu(A) \, | \, \mu \in M_{\ens}\})$.\footnote{TODO: prove somewhere}
+
+As we saw, the classical case corresponds to a simplex in Choquet theory.
+This means that each element of the ensemble space is represented by a unique probability measure over the extreme points.
+In that case, the fraction capacity extended on the extreme points will also be additive.
+In the non-classical case, the closure of the ensemble space will not be a simplex, and there will be multiple representations, which will lead to a non-additive measure.
+Suppose, in fact, that $\mu$ and $\nu$ are two distinct measures that represent the same ensemble $\ens$.
+Then there will be a set $A$ such such that $\mu(A) \neq \nu(A)$.
+Suppose, without loss of generality, that $\mu(A) > \nu(A)$ We will also have $\mu(A^{\complement}) = 1- \mu(A) < 1 - \nu(A) = \nu(A^{\complement})$.
+Since the fraction capacity for a set cannot be less than the measure on that set, we have $\frcap_{\ens}(A \cup A^{\complement})=\frcap_{\ens}(X) = 1 = \mu(A) + \mu(A^{\complement}) < \mu(A) + \nu(A^{\complement}) \leq \frcap_{\ens}(A) + \frcap_{\ens}(A^{\complement})$.
+That is, the fraction capacity is additive over disjoint sets of extreme points in and only in the classical case.
+In other words, the fraction capacity reduces to a classical probability distribution over pure states in the classical case.
 
@@
 
-\section{Fraction capacity}
-
-Given that mixing coefficients have a more general applicability than probability of outcomes, let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
-How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
-For example, let $\Ens$ be the space of all probability distributions for a six-sided die.
-Let $\ens_{123456}$ be the uniform distribution over all outcomes.
-Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
-Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$, where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/6$ of $\ens$ can be constructed from $A$, but no more.
-Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
-Note that, given the uniform distribution, $1/6$ is exactly the probably for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
-This gives the basic insight for our definition.
-
-Given a target ensemble $\ens \in \Ens$ and an arbitrary ensemble $\ens[a] \in \Ens$, we define the \textbf{fraction} of $\ens[a]$ in $\ens$ to be
-\begin{equation}
-	\fraction_{\ens}(\ens[a]) = \sup(\{ p \in [0,1] \, | \, \exists \, \ens_1 \in \Ens \text{ s.t. }  \ens = p \ens[a] + \bar{p} \ens_1 \}).
-\end{equation}
-The fraction is always well-defined because we can always write $\ens = 0 \ens[a] + 1 \ens$, therefore it must be zero or greater.
-This quantity tells us how much of ensemble $\ens$ can be constructed from $\ens[a]$.
-
-We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a Borel set of ensembles $A \subset \Ens$.
-Recalling that the $\hull(A)$ is the set of all possible convex combinations (i.e.
-mixtures), we define the\textbf{fraction capacity} of $A$ for $\ens$ to be
-\begin{equation}
-	\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\}).
-\end{equation}
-This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$.\footnote{The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
-fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.
-TODO \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} }
-
-One can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
-\begin{enumerate}
-	\item non-negative and unit bounded - $\frcap_{\ens}(A) \in [0,1]$
-	\item monotone - $A \subseteq B \implies \frcap_{\ens}(A) \leq \frcap_{\ens}(B)$
-	\item sub-additive - $\frcap_{\ens}(A \cup B) \leq \frcap_{\ens}(A) + \frcap_{\ens}(B)$
-	\item continuous from below and above - $\frcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \frcap_{\ens}(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
-\end{enumerate}
-
-The first property readily comes from the domain of the fraction.
-For the second, note that $\hull$ and $\sup$ are both monotone.
-For sub-additivity, note that the fraction of two components can at most sum during mixing.
-For the last, increasing and decreasing sequences strictly add or remove possible mixtures, therefore they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
-The limit of these sequences must agree with the fraction capacity of the limit.
-
-If not for the lack of additivity over disjoint sets, the fraction capacity would be a probability measure.
-However, the difference between fraction capacity and probability goes beyond that.
-While the probability measure is defined over sets of outcomes, the fraction capacity is defined over ensembles, over probability distributions.
-Conceptually the two are different object, but they, as we will see later, linked.
-Suppose $\Ens$ is the classical ensemble space given by the set of probability distributions over a sample space $\Omega$.
-Let $U \subseteq \Omega$ be an event, let $A_U \subset \Ens$ be the set of all probability distributions with support within $U$ and $\ens \in \Ens$ a probability distribution.
-Then we can show that $\frcap_{\ens}() $
-
-In the classical case, the fraction capacity reduces to a probability measure.
-Note that a probability satisfies all the properties of a fraction capacity, but it also satisfy on additional property: additivity over disjoint sets.
-That is, if $A \cup B = \emptyset$ then $\mu(A \cup B) = \mu(A) + \mu(B)$.
-Adding finite additivity over disjoint sets is enough as continuity will extended the property to additivity over countably many disjoint sets.
-Conversely, having countable additivity over disjoint sets implies sub-additivity and continuity from below and above.
-
-While additivity over disjoint set holds in classical probability, it does not hold in quantum mechanics
-
- This is not true in general, as we can see in a simple example from quantum mechanics.
-Let $\Ens$ be the ensemble space of a qubit, that is the space of density matrices of a two dimensional Hilbert space.
-Geometrically, this is the convex set represented by a three dimensional ball.
-Let $\ens_{\psi} = |\psi\>\<\psi|$ be a pure state and let $\ens_{\phi} = |\phi\>\<\phi|$ be its orthogonal.
-Geometrically, these are two opposite points on the surface of the ball.
-The maximally mixed state can be expressed as $\ens = \frac{1}{2} \ens_0 + \frac{1}{2} \ens_1$.
-This is the center point of the ball and is in fact the midpoint between two opposite points.
-Since this is valid for any point, we have $\frcap_{\ens}(\{\ens_{\psi}\})=\frac{1}{2}$ for any $\ens_{\psi} \in \Ens$ and $\frcap_{\ens}(\{\ens_{\psi}, \ens_{\phi}\})=1$ for any pair of orthogonal pure states.
-If $A$ is a set of two orthogonal pure states, and $B$ is state different from the two, we have $\frcap_{\ens}(A \cup B) = 1 < \frcap_{\ens}(A) + \frcap_{\ens}(B) = \frac{3}{2}$.
-
-If we restrict ourselves to lattice of subsets of $\Ens$ we can recover additivity.
-For example, if we take a family $A_i$ of orthogonal sets (i.e.
-all the elements of one set are orthogonal to all the elements of any other set) and take an element that is a mixture of elements from that family, then the fraction capacity will be additive over the lattice formed by closing the family under union and intersection.
-In quantum mechanics, this would correspond to all probability distributions that are outcomes of the projective measurement that distinguishes between $A_i$.
-Part of the task, then, will be finding a set of necessary and sufficient conditions under which the fraction capacity for a particular ensemble is additive over a particular lattice of subsets
-
-What is important here is that
-
 \section{Statistical properties}

### 2025-05-13 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: removing old stuff

@@
 
-\section{Old introduction}
-
-The goal of this paper is to show that, in the same way that classical probability can be axiomatized through standard measure theory, quantum probability, and in fact any probability used a physical theory, can be axiomatized through non-additive (i.e.
-fuzzy) measure theory.
-We will define a generalized concept of probability that is ultimately based on physically motivated axioms, giving the physicist a clear operational motivation to our definitions, the mathematicians a formally well-defined mathematical structure and the philosopher a unified framework in which probability can be interpreted in the context of all physical theories.
-We are not going to be able to develop a full theory here, as there are still a number of conceptual and technical challenges to be solved that will require work in physics, mathematics and philosophy.
-However, we will show that there are enough results that point in the same direction, and therefore make it reasonable to invest time and effort to work on the open problems.
-
-The material presented is part of a larger work that aims to coalesce ideas and results from different branch of mathematics and physics.
-For those interested, the current state of the research is openly available in an open document that will be routinely updated as experts from different fields review and help us improve the overall theory.
-
-In a nutshell, we will start by arguing that physical theory must at least be able to talk about statistical ensembles, and that ensembles must allow the preparation of statistical mixture.
-Given a target ensemble $\ens$ and a set of ensemble $A$, we define the fraction capacity of $A$ for $\ens$ as the biggest fraction of $\ens$ that can be constructed from another set of ensembles $A$.
-That is, the highest $p$ such that we can express our ensemble as $\ens = p \ens[a] + (1-p) \ens[b]$, where $\ens[a]$ is a mixture of elements of $A$.
-For example, if $\ens$ is the uniform distributions for a roll of a six-faced die, and $A$ includes only two ensembles with $100\%$ probability for the outcomes $3$ and $4$ respectively, then the faction capacity of $A$ for $\ens$ will be $\frac{1}{3}$.
-Note that this coincides with the probability of the event ``$3$ or $4$''.
-The fraction capacity is a non-negative, unit bounded, sub-additive and continuous set function, which is therefore like a probability measure except for the additivity, which is recovered for classical mechanics and over quantum measurements.
-
-Similarly, given a set $A$ of ensembles we define the \textbf{state capacity} as the exponential of the highest entropy obtainable with a mixture of $A$.
-Note that the highest entropy of a subspace in both classical and quantum statistical mechanics is given by the uniform distribution, therefore the state capacity recovers the count of states (i.e.
-the Liouville volume) in classical mechanics and the dimensionality of the Hilbert space in quantum mechanics.
-The state capacity is also a non-negative, sub-additive and continuous set function.
-This points to a possible generalization of measure theoretic calculus to a sub-additive case that would be guaranteed to work on all physical theories.
-
-\section{Ensemble spaces}
-
-Note that our goal is not to construct an abstract probability theory that applies to everything, but rather to construct a specific probability theory that is guaranteed to apply to all physical systems.
-Therefore we do not to define an axiomatic theory of probability and apply it to physical systems, but rather characterize basic properties of physical states and from those build a theory of probability that fits those definitions.
-Whether it is applicable to decision theory, economics, or psychology is out of scope.
-As is common in our larger project, we start from axioms that, in our opinion, represent requirements of scientific practice.
-We call these \textbf{constitutive assumptions} as they are essential to proceed with the scientific endeavor.
-The advantage is that the realm of applicability and physical interpretation will be clear from the start.
-
-We start from requiring a physical theory to allow statistical descriptions: at the very least, it will provide a notion of ensemble and describe which ensembles that are allowed by the theory.
-An ensemble can be understood as the infinite collection of the outputs of a particular preparation procedure.
-The ensemble space of classical mechanics, for example, is the space of probability distributions on phase space.
-In quantum mechanics, instead, it is the space of density operators.
-While the ensemble space for each theory will in principle be different, they will all share common features: they must allow measurable properties of the system, probability distributions for outcomes of said properties, characterize reversible and irreversible processes through an entropy function, and so on.
-We are looking for necessary minimal basic requirements upon which these common tools can be built.
-
-At least two arguments can be made to show that any physical theory must allow statistical descriptions, i.e.~provide an ensemble space.
-First, note that physical laws are never about specific instances but rather regularities.
-They are statements of the type ``whenever we prepare this, we can measure that.'' This means that, ultimately, they are about ensembles.
-Second, if a physical theory is to be testable experimentally, and repeatedly so, it must be in terms of ensembles because this is what we prepare in practice and this is what, in the end, we characterize experimentally.
-Moreover, repeatability requires the ability to always test ``one more time,'' which implies that ensembles are infinite collections, and therefore idealization.
-The approach, then, can most likely be given firm conceptual grounds, which we leave as an open but likely solvable philosophical problem.
-
-We have identified three basic requirements on ensembles: first, they have to be identifiable experimentally; second, they need to allow statistical mixtures; lastly, they need a well-defined entropy.
-Let us go through these items one by one.
-We leave the full details to the working document (TODO: decide how to call it).
-
-\subsection{Experimental verifiability}
-The first requirement is that ensembles must be connected to experimental verification.
-That is, we must have enough \textbf{verifiable statements} at our disposal to define ensembles and tell them apart.
-By verifiable statement we mean a statement for which an experimental test is available that terminates in finite time if and only if the statement is true.
-For example, the statement ``the mass of the photon is less than $10^{-13} \, eV$'' is verifiable, while ``the mass of the photon is exactly $0 \, eV$'' is not verifiable due to its infinite precision.
-Based on previous work, (CITE) this mathematically requires an ensemble space to be a $T_0$ second-countable topological space, where open sets represent verifiable statements and Borel sets represent statements associated with a test, but without guarantee of termination.
-
-The topological structure, then, represents the most foundational structure for a physical theory.
-It tells us why functions are ``well-behaved,'' which simply means they have to be topologically continuous in the ``natural topology'' induced by this requirement.
-It tells why probability is assigned to Borel sets (i.e.
-all statements associated with tests) and why the Banach-Tarski paradox does not apply in physics (i.e.
-non-Borel sets are physically ill-defined as they are not connected to experimental verification).
-Similarly, it tells us that sets with cardinality greater than that of the continuum are not physically relevant, as they cannot be given a $T_0$ second-countable topology.
-
-In the case of ensembles, typical verifiable statements will be ``the average energy of the particle is between 3 and 4 $eV$,'' ``the probability of getting heads is between 49 and 51 percent,'' ``the probability distribution for the position is a Gaussian of mean 0 $m$ and 1 $m$ standard deviation within 1\%.'' Note that these are not results of single-shot measurements, and therefore verifiable statements are distinctly different from observables in quantum mechanics.
-In the same manner, these are not measurements that extract one bit of information, as there is no requirement of termination in the negative case.
-
-\begin{tcolorbox}[colback=white, colframe=black]
-	Experimental verifiability $\Rightarrow$ An ensemble space $\Ens$ is a $T_0$ second-countable topological space.
-\end{tcolorbox}
-
-\subsection{Statistical mixtures}
-Another basic requirement is the ability to prepare a \textbf{mixture} of two ensembles.
-If $\ens[a], \ens[b] \in \Ens$ are two ensembles, and $p \in [0,1]$ a weight, then $\ens = p \ens[a] + (1-p) \ens[b] \in \Ens$ is the ensemble that describes a process that selects the first ensemble over the second $p$ percent of the times.
-Mathematically, the ensemble space is endowed with a \textbf{convex structure}.
-Moreover, since the mixing operation must be consistent with experimental verifiability, it will be topologically continuous, and the ensemble space will be a topological convex space.
-
-Ultimately, the convex structure is responsible for all linear structures we have in physics.
-In face, the entropy and its bounds, defined later, forces the ensemble space to be the subset of a convex subset of a vector space.
-That is, the convex structure is ``invertible.''\footnote{If we fix $\ens$, $\ens[a]$ and $p$ then, if it exists, there is only one $\ens[b]$ such that $\ens = p \ens[a] + (1-p) \ens[b]$.
-In the context of convex spaces, this property is called ``cancellative.''} What is still not clear is whether the ensemble space embeds continuously in a topological vector space.
-While topological vector spaces are well established in the literature, topological convex spaces are not.
-
-Regardless of the topology, we are already equipped to understand what negative probability is.
-Suppose we have a real vector space of dimensions $n$, then $n$ linearly independent vectors are enough to express every point as a linear combination.
-We can also take an additional vector and express every point as an affine combination, that is a linear combination where the coefficients sum to one, though some can still be negative.
-Since every ensemble space, including the space of mixed states in quantum mechanics, is a subset of a real vector space, we can choose a linearly dependent set of states that span the whole space, and every mixed state can be expressed as an affine combination, as a pseudo-probability.
-However, the boundaries of the ensemble space will determine which affine combinations are valid states (i.e.
-they end up within the ensemble space), and there is no requirement that these boundary are conveniently expressible in terms of affine combinations.
-
-The convex structure, however, allows us to ask whether two ensembles have a \textbf{common component}.
-That is, then can be seen as different statistical mixture that have some part in common.
-Two ensemble are separate, noted $\ens[a] \separate \ens[b]$, if they do not have a common component.
-This allows to characterize the space in terms of statistical decomposition, and see how this differs from other decompositions.
-
-\begin{tcolorbox}[colback=white, colframe=black]
-	Statistical mixture of ensembles $\Rightarrow$ An ensemble space $\Ens$ is a topological convex space.
-\end{tcolorbox}
-
-\subsection{Entropy}
-The last requirement is that each ensemble must have a well defined \textbf{entropy}. That is, there is a scalar function $S : \Ens \to \mathbb{R}$ that satisfies the following:
-\begin{enumerate}
-	\item strictly concave: $S(p \ens[a] + (1-p) \ens[b]) - (p S(\ens[a]) + (1-p) S(\ens[b]) ) \geq 0$
-	\item bounded from above: $S(p \ens[a] + (1-p) \ens[b]) - (p S(\ens[a]) + (1-p) S(\ens[b]) ) \leq - p \log p - (1-p) \log(1-p)$
-\end{enumerate}
-The first tells us that, during mixing, the entropy cannot decrease, and it stays the same if and only we are mixing an ensemble with itself.
-The second tells us that the most the entropy can increase is given by the choice between the two ensembles.\footnote{The expression for the Shannon entropy is actually derived instead of imposed axiomatically.} We define two ensembles to be \textbf{orthogonal} when they maximally increase the entropy during mixing.
-This recovers the standard notion of orthogonality: two classical ensembles are orthogonal when they have disjoint support, and therefore $\int_M \rho_1 \rho_2 d\mu = $; two quantum ensembles are orthogonal when they are defined on orthogonal subspaces, and therefore $\tr(\rho_1 \rho_2) = 0$.
-
-The entropy is ultimately responsible for all geometrical structure in physics.
-In classical mechanics, the geometry is essentially defined by phase space volumes and areas in each degree of freedom (DOF).
-The entropy of uniform distributions is given by the logarithm of the volume, meaning that given the volume we are able to calculate the entropy and given the entropy we are able to reconstruct the volume.
-Similarly, uniform marginals will recover DOF areas.
-In quantum mechanics, the inner product (i.e.
-the Born rule) allows us to calculate the entropy of the mixture of two pure states, and given the entropy we can recover the Born rule.
-
-An interesting insight is that the entropy, being strictly concave, has a negative defined Hessian.
-The negation of the Hessian, then, is a symmetric positive defined function of two infinitesimal variations, and can serve as a metric tensor on the affine structure given by the mixing coefficients.
-In both classical and quantum mechanics this recovers the Fisher-Rao metric.
-
-Another crucial feature is that two ensembles maximize the entropy increase if and only if they are orthogonal: two classical ensembles maximize the entropy increase when they have disjoint support, and therefore $\int_M \rho_1 \rho_2 d\mu = 0$; two quantum ensembles when they are defined on orthogonal subspaces, and therefore $\tr(\rho_1 \rho_2) = 0$.
-This corresponds to the case when the two ensemble are made of instances that are mutually exclusive.
-Therefore we define two ensembles to be \textbf{orthogonal}, noted $\ens[a] \ortho \ens[b]$, if their mixture leads to a maximal entropy increase.
-The following additional axiom is valid for orthogonal ensembles
-\begin{enumerate}
-	\setcounter{enumi}{2}
-	\item mixtures preserve orthogonality: $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$
-\end{enumerate}
-as mixing elements distinguishable from those of ensemble $\ens[a]$, still gives elements distinguishable from ensemble $\ens[a]$.
-
-It can be shows that orthogonal ensembles are separate, but the converse is not true.
-In quantum mechanics, for example, all pure states are separate (they cannot be decomposed and therefore cannot have components in common) but not all are orthogonal.
-In classical spaces, however, all separate ensembles are also orthogonal.
-Understanding the relationships between these two properties, then, is key in understanding the difference between classical and quantum mechanics, and what can happen in other physical theories.
-
-\begin{tcolorbox}[colback=white, colframe=black]
-	Existence of entropy $\Rightarrow$ Strictly concave function whose upper bound defines orthogonality. Statistical mixtures preserve orthogonality.
-\end{tcolorbox}
-
-\subsection{Interplay between structures}
-
-Before introducing the actual measures, we want to note that there is a delicate balance between these structures which is important at both a conceptual level and at a more mathematical level.
-Part of the difficulty is to recognize which standard modes of understanding the problem should be kept, as the they generalize nicely, and which should be discarded.
-Let us go through some key examples.
-
-Those that focus on the convex/linear structure may be tempted to close the space under infinite convex/linear combinations.
-It would be very convenient mathematically, and, in fact, this is what happens in Hilbert spaces.
-However, this closure leads yo physically ill-defined systems, such as a particle with an infinite average position or undefined energy.(CITE Hilbert) It is the connection to experimental verification, the topology, that will tell us which limits are allowed and which are not.
-Therefore, the lack of infinite convex is a crucial feature of the space, and convex sets in ensemble spaces are to be understood as closed convex sets, so that closure on the topology includes those, and only those, infinite mixtures that are physically meaningful.
-
-If we focus on probability, we are tempted to extend the intuition from classical mechanics and it is useful to understand exactly why and where it fails.
-On a convex structure we can define the extreme points, those that cannot be further decomposed.
-Therefore it is tempting to define \textbf{pure states} as those ensembles that cannot be further decomposed into other ensembles.
-For example, the pure states for a six-sided die are the ensembles perfectly prepared respectively on the six possible outcomes.
-All ensembles are convex combinations $\sum_{i=1}^{6} p_i \ens_i = 1$ of those pure states $\ens_i$.
-This is exactly how classical discrete spaces work and much of our intuition for probability comes from there.
-
-This approach, however, already becomes problematic in classical mechanics.
-Since we can take uniform distributions over any region of phase space, we can imagine to take a sequence of more and more refined distributions around a value of position and momentum.
-Therefore we may be tempted to say that the points of phase space represent perfectly prepared systems, and every other probability is a distribution on these space.
-This is physically unsound.
-First of all, we cannot consistently prepare position and momentum with infinite precision, therefore those ensembles do not exist even exist as an idealization.
-Secondly, the entropy of those ensembles, if they existed, would be minus infinity, which is a problem for any thermodynamic treatment.
-To make this intuition consistent with the math would require each point in phase space to be topologically isolated, which would extend the $\sigma$-algebra to the power set of $\mathbb{R}^n$, which means there would be no consistent way to define a volume.
-The fact that the math breaks down is simply a symptom that our physical problem is ill-specified.
-
-In quantum mechanics, things are even worse.
-Pure states are indeed extreme points of the convex set.
-However, we also have discrete spectra.
-Here the temptation is, like we ``did'' in classical mechanics, to consider wave-functions all concentrated at a single value of position.
-While one may think this is the analogue of classical mechanics, it is not.
-We can, in fact, take a wave-function that is uniformly distributed on every finite range of position, and we can imagine to shrink that range.
-However, this is a sequence of pure states.
-All pure states have zero entropy, and the reduced spread on position will mean an increased spread on momentum.
-That is, we are not providing a sequence of ensembles at ever greater prevision, like in the classical case.
-In other words, the eigenstates of position in quantum mechanics, even if we wanted to say they were actual states, they would not be the limit of infinitesimal decomposition (i.e.
-a limit in the convex structure) but rather the limit of a path over pure states (i.e.
-a topological limit).
-
-The key insight is that \textbf{mixing coefficients and probabilities are different in general} and they coincides only when a particular ensemble can be understood as a statistical mixtures of perfect preparations of a given quantity (i.e.
-of orthogonal ensembles).
-This always happens in the classical case and over measurement contexts in the quantum case, but is not the general case.
-For example, an equal mixture of spin up and spin left, for example, does not correspond to a 50\% probability of measuring spin up.
-The key to develop a general theory of probability, then, is to characterize the mixing coefficients first and understand when they recover probabilities of outcomes.
-
-To recap, the topological structure captures experimental verifiability and is responsible for the limits and the connection to measureable quantities.
-The convex structure captures statistical mixing and is responsible for all linear structures of physical theories.
-The entropic structure captures the variability within an ensemble and is ultimately responsible for all geometric structures.
-Since the physical problem is indivisible, there is an interplay across these structure which goes beyond this article and we refer to our larger work.
-
-\section{Statistical properties}
-
-\section{Beyond real valued quantities}
-
-\section{State capacity}
-
-\section{Quantization}
-* 3 pick 2
-
-\section{Quantizing space-time}
-* we need a non-additive measure on degrees of freedom
-*
-
 \section{Conclusion}

### 2025-05-25 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: adding state capacity section

@@
 
+\newcommand\mix{\mathrm{mix}}
+\newcommand\component{\mathrm{comp}}
+\newcommand\cospan{\mathrm{cospan}}
+\newcommand\dist{\mathrm{dist}}
 \newcommand\hull{\mathrm{hull}}
-\newcommand\stcap{\mathrm{scap}}
+\newcommand\support{\mathrm{supp}}
+\newcommand\capacity{\mathrm{scap}}
 \newcommand\fraction{\mathrm{frac}}
@@
 
-The first property readily comes from the domain of the fraction.
+The first property readily comes from the range of the fraction.
 For the second, note that $\hull$ and $\sup$ are both monotone.
@@
 
+\section{State capacity}
+
+We now want to introduce a measure theoretic representation of the entropy that generalizes the notion of count of states.
+Since entropy characterizes the variability of the preparations of the ensemble, greater variability means the ensemble is spread over more cases.
+Therefore the count of distinguishable states must be a monotonic function of the entropy.
+It has been shown[cite Hall] under very general conditions that the exponential of the entropy is the only function to have the appropriate features.
+One required feature, that we prove in general, is that the exponential of the entropy is subadditive (i.e.
+the count of states associated to a mixture cannot exceed the combined count of states associated to both mixtures) and it is additive for orthogonal ensembles (i.e.
+no overlap in possible instances) in a specific linear combination.
+
+Given a Borel set of ensembles $A \subseteq \Ens$, we define the \textbf{state capacity} of $A$ to be
+\begin{equation}
+\capacity(A) = \sup(2^{S(\hull(A))}\cup\{0\}).
+\end{equation}
+This returns the exponential of the biggest entropy achievable with a mixture of elements of $A$.
+
+One can then show that the fraction capacity $\capacity_{\ens} : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
+\begin{enumerate}
+	\item non-negative - $\capacity(A)\geq 0$
+	\item monotone - $A \subseteq B \implies \capacity(A) \leq \capacity(B)$
+	\item sub-additive - $\capacity(A \cup B) \leq \capacity(A) + \capacity(B)$
+	\item additive over orthogonal sets - $\capacity(A \cup B) = \capacity(A) + \capacity(B)$ if $A \ortho B$
+	\item continuous from below and above - $\capacity(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \capacity(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
+\end{enumerate}
+
+The first property readily comes from the range of the exponential.
+For the second, note that $\hull$, exponential and $\sup$ are all monotone.
+For sub-additivity, the exponential of the entropy itself is sub-additive, and it is additive over orthogonal ensembles.
+For the last, since the entropy itself  is continuous, this property carries over to the set function.
+
+Again we note that the non-additive measure we recover is very well-behaved, and connects directly with important quantities.
+When Choquet theory applies, state capacity can be extended to the extreme points.
+In the classical case, disjoint sets of extreme points will be orthogonal, meaning that the state capacity there restricted is an additive measure.
+In the discrete case, every point is associated to zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
+In the continuous case, gives a subset of phase space $A$, the maximum entropy is reached by the uniform distribution $\rho_A$, which means we recover $\log_2 \capacity(A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
+Similarly, in quantum mechanics, given an orthogonal subspace $A$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(A) = 2^{\dim(A)}$.
+
+The state capacity, then, recovers the number of distinguishable states.
+In classical mechanics, all states are perfectly distinguishable.
+In quantum mechanics, only orthogonal state are distinguishable, and therefore the state capacity recovers the number of orthogonal states.
+
+\section{Connection to fuzzy logic}
+
+\section{Possible generalization of calculus}
+
+As we pointed out, the state capacity and the fraction capacities are always well-defined, even in the cases where, potentially, Choquet theory does not apply because the closure of the ensemble space is not compact.
+
+General problems:
+1) Can we always find extreme points?
+2) Can we generalize to the non-compact case? Can \href{https://en.wikipedia.org/wiki/Krein%E2%80%93Milman_theorem#convex_compactness}{convex compactness} help?
+3) In the classical case, both measures become additive on the extreme points. Standard calculus is used. Is there a non-additive version of calculus that generalizes?
+
 \section{Conclusion}

### 2025-06-04 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Typos

@@
 For the more mathematically inclined, we will point to areas where new math or generalization of established math are needed.
-For the more philosophically incline, we will point to areas where further conceptual work is needed.
+For the more philosophically inclined, we will point to areas where further conceptual work is needed.
 
@@
 Both are non-negative sub-additive monotonic set functions, with the second being unit bounded.
-These measures are always unique and, since they are defined on the ensemble space itself, always have provide a physically meaningful representation.
-Moreover, in the compact case, these measure can be restricted on the extreme points while retaining uniqueness.
+These measures are always unique and, since they are defined on the ensemble space itself, always provide a physically meaningful representation.
+Moreover, in the compact case, these measures can be restricted to the extreme points while retaining uniqueness.
 (TODO: prove)
@@
 
-Additional, each ensemble space $\Ens$ is equipped with a continuous entropy function $S : \Ens \to \mathbb{R}$. The entropy of a statistical mixture satisfies the following bounds
+Additionally, each ensemble space $\Ens$ is equipped with a continuous entropy function $S : \Ens \to \mathbb{R}$. The entropy of a statistical mixture satisfies the following bounds
 \begin{equation}
@@
 Intuitively, the variability cannot decrease during mixing, and will have an upper bound when mixing two ensembles that are non-overlapping, or orthogonal.
-In fact, we define two ensembles $\ens[a] \ortho \ens[b]$ to be orthogonal when they mixture gives a maximal entropy increase.
+In fact, we define two ensembles $\ens[a] \ortho \ens[b]$ to be orthogonal when their mixture gives a maximal entropy increase.
 This does recover the notion of orthogonality in the corresponding classical and quantum vector spaces of statistical ensembles.
@@
 That is, all statistical properties of $\ens$ can be recovered by the probability measure.
-This is what allows to think go $\ens$ as a probability distribution.
+This is what allows us to think of $\ens$ as a probability distribution.
 Note, however, that the Choquet theorem does not guarantee uniqueness.
@@
 
-We can apply Choquet theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
+We can apply Choquet's theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
 TODO: finalize the proof} and metrizable.
-In this case, we are guaranteed that all ensembles allows a representation through a probability on points that may corresponds idealized ensembles (i.e.
+In this case, we are guaranteed that all ensembles allow a representation through a probability on points that may correspond to idealized ensembles (i.e.
 the Dirac measure) that are not, even conceptually, realizable.
@@
 
-Chouqet theory tells, in the realm where it can be applied, that classical probability fails in quantum mechanics precisely because the ensemble space of a quantum system allows multiple decomposition.
+Chouqet theory tells us, in the realm where it can be applied, that classical probability fails in quantum mechanics precisely because the ensemble space of a quantum system allows multiple decompositions.
 
@@
 Another way to represent quantum ensembles is through the use of quasiprobability distributions.
-Given a space $X$, a measure $\mu$ is affine if $\mu(X)$.
-All probability measures are affine measures, but the reverse is not true as affine measures can be negative on particular regions.
+Given a measurable space $(X, \Omega_X)$, a measure $\mu : \Omega_X \to \mathbb{R}$ is affine if $\mu(X) = 1$.
+All probability measures are affine measures, but the reverse is not true, as affine measures can be negative on particular regions.
 The Wigner function, for example, can be negative over small regions of phase space.
@@
 
-In general, since the ensemble space is bounded along every direction, not all affine measure will correspond to an ensemble space.
-Whether all probability measure correspond to a valid ensemble, or whether affine measures are even used, correspond to relationships between $\Ens$ and the space of probability measures $M_p(X)$.
+In general, since the ensemble space is bounded along every direction, not all affine measures will correspond to an ensemble space.
+Whether all probability measures correspond to a valid ensemble, or whether affine measures are even used, corresponds to relationships between $\Ens$ and the space of probability measures $M_p(X)$.
 If $\Ens \subset M_p(X)$, then all states are represented by a probability measure.
 However, this is not a ``probability'' in the standard sense of the word.
-For example, the Husimi Q representation is always non-negative at every point, but two points to do not represent mutually exclusive events as they are not orthogonal (i.e.
+For example, the Husimi Q representation is always non-negative at every point, but two points do not represent mutually exclusive events as they are not orthogonal (i.e.
 their inner product is non-zero and they do not maximize the entropy increase during mixture).
-Mathematically, the issue is that the points $X$ of the Husimi Q representation are not the extreme points of the closure ensembles space $\overline{\Ens}$.
+Mathematically, the issue is that the points $X$ of the Husimi Q representation are not the extreme points of the closure ensemble space $\overline{\Ens}$.
 In general, $\Ens$ and $M_p(X)$ will overlap, but neither will contain the other.
-For example, in the Wigner representation the only pure states (i.e.
+For example, in the Wigner representation, the only pure states (i.e.
 the extreme points of $\Ens$) represented by a probability measure are the Gaussian wave-packets.

### 2025-06-04 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: minor changes

--- 2025-05-25 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 These mathematical properties are physically grounded in the ability to experimentally define ensemble (leading to a topological space), to define statistical mixtures (leading to a convex space), to quantify the variability of the elements of the ensembles (leading to an entropy function) and to having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
-The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the number of orthogonal subspaces of an operator)?
+The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the dimension of orthogonal subspaces of an operator)?
 What measure theoretic tool can represent each ensemble?
@@
 
+[This section has to make clear in what extra conditions on ensembles spaces Choquet theory can be applied.]
+
+% Intro paragraph that says quickly how to connect to Choquet theory and its limitations. It should be a quick summary of the whole section
 Since ensemble spaces are a convex subset of a locally convex topological vector space, they can be connected to Choquet theory.
+
+% Quick review of Choquet theory
+Let us first give a summary of the main results in Choquet theory that are interesting for our purposes.
 First of all, let $E$ be a convex subset of a locally convex space $V$.
-We say that $\ens \in E$ is represented by a probability measure $\mu : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_X f d\mu$ for every continuous linear functional $f$ on $V$.
+We say that $\ens \in E$ is represented by a probability measure $\mu : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_E f d\mu$ for every continuous linear functional $f$ on $V$.
 The Choquet theorem states:
@@
 
+% Physical meaning of representation with a measure.
 First, let us understand the physical significance.
@@
 
+% Limitation of applying Choquet theory to our case
 We should now look at the requirements of the theorem, to see whether we can always apply it to an ensemble space.
@@
 
+% Define "Choquet space" as those spaces where we can apply Choquet theory
 We can apply Choquet theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
@@
 Another way to represent quantum ensembles is through the use of quasiprobability distributions.
-Given a space $X$, a measure $\mu$ is affine if $\mu(X)$.
+Given a topological space $X$, a measure $\mu$ is affine if $\mu(X)=1$.
 All probability measures are affine measures, but the reverse is not true as affine measures can be negative on particular regions.
@@
 To develop a general theory for physics, however, this is not useful as each representation is not general enough.
+
+TODO: disjoint supports of Wigner functions do not give orthogonal subspaces (CHECK)
 

### 2025-06-06 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex

--- 2025-06-04 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 That is, all statistical properties of $\ens$ can be recovered by the probability measure.
-This is what allows us to think of $\ens$ as a probability distribution.
+This is what allows to think go $\ens$ as a probability distribution.
 Note, however, that the Choquet theorem does not guarantee uniqueness.
@@
 
+% Define "Choquet space" as those spaces where we can apply Choquet theory
 We can apply Choquet's theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
@@
 Another way to represent quantum ensembles is through the use of quasiprobability distributions.
-Given a measurable space $(X, \Omega_X)$, a measure $\mu : \Omega_X \to \mathbb{R}$ is affine if $\mu(X) = 1$.
-All probability measures are affine measures, but the reverse is not true, as affine measures can be negative on particular regions.
+Given a topological space $X$, a measure $\mu: \mathcal{B}(X) \to \mathbb{R}$ is affine if $\mu(X)=1$.
+All probability measures are affine measures, but the reverse is not true as affine measures can be negative on particular regions.
 The Wigner function, for example, can be negative over small regions of phase space.

### 2025-06-10 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Fixed typos and grammar
using Grammarly

@@
 \begin{abstract}
-	We present a physically motivated generalization of probability theory that is suitable for classical mechanics, quantum mechanics and any future physical theory that allows a statistical description.
-The goal is to put the use of classical and quantum probability in a broader context, and show how the current mathematical structures are likely not suitable to solve the open problems in the foundations of physics.
-For the more mathematically inclined, we will point to areas where new math or generalization of established math are needed.
+	We present a physically motivated generalization of probability theory that is suitable for classical mechanics, quantum mechanics, and any future physical theory that allows a statistical description.
+The goal is to put the use of classical and quantum probability in a broader context and show how the current mathematical structures are likely not suitable to solve the open problems in the foundations of physics.
+For the more mathematically inclined, we will point to areas where new math or generalization of established math is needed.
 For the more philosophically inclined, we will point to areas where further conceptual work is needed.
@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
+A key problem, then, is what notion of probability should be used, given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
 
 We reformulate the problem in the following way: given the space of all possible statistical ensembles allowed by a statistical theory, what measure theoretic tools can we use to represent each ensemble and the geometric properties of the overall space?
-Why exactly classical probability measures fail to capture completely quantum states?
+Why exactly do classical probability measures fail to capture completely quantum states?
 What suitable generalizations are possible, and what mathematical tools would need to be developed?
@@
 These mathematical properties are physically grounded in the ability to experimentally define ensemble (leading to a topological space), to define statistical mixtures (leading to a convex space), to quantify the variability of the elements of the ensembles (leading to an entropy function) and to having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
-The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the dimension of orthogonal subspaces of an operator)?
+The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurement (as in the dimension of orthogonal subspaces of an operator)?
 What measure theoretic tool can represent each ensemble?
@@
 We will first review some results from Choquet theory, which is, to our knowledge, the most general approach to represent elements of a convex set using additive probability measures.
-The key insight is that, at least in the compact case, such representation is always possible, but it is not unique.
+The key insight is that, at least in the compact case, such a representation is always possible, but it is not unique.
 Uniqueness requires the space to be a Choquet simplex, which is a generalization of the standard finite-dimensional simplex and is a particularly shaped subset of a vector space.
@@
 That is, there is no unique or privileged choice for this representation.
-Moreover, since the map is at the level of the full vector space, not just the subset with valid ensembles, there is no general way to characterize which subset of signed probability measures correspond to ensembles.
+Moreover, since the map is at the level of the full vector space, not just the subset with valid ensembles, there is no general way to characterize which subset of signed probability measures corresponds to ensembles.
 Therefore, while signed measures are a powerful tool for calculation in specific problems, they cannot give us a general principled representation of the physical theory.
@@
 The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
-From the entropy, we can define the state capacity, which generalizes the notion of count of states.
+From the entropy, we can define the state capacity, which generalizes the notion of the count of states.
 For each ensemble, we can define the fraction capacity, which generalizes the notion of probability.
@@
 The fraction capacity is the supremum of all possible measures for the same ensemble.
-The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure, and is therefore additive.
+The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure and is therefore additive.
 The state capacity, instead, corresponds to the minimum of the entropy of all possible representations.
@@
 Having shown that a general approach based on non-additive measures is possible, we lay down a series of questions to explore in the future.
-The biggest one is whether it is possible to generalize notions such as integration, derivation and expectation to this non-additive case.
-It should be noted that non-additive measures are used in the context of decision theory, so there may be a connection between the use of non-additive measures in quantum theory and interdependent complimentary choices.[TODO: find citation that sub-additive is when choices are complementary] Both the Choquet and the Sugeno integral, for example, are defined over a monotonic set-functions, so there is a chance that these tools may turn out to be useful.
+The biggest one is whether it is possible to generalize notions such as integration, derivation, and expectation to this non-additive case.
+It should be noted that non-additive measures are used in the context of decision theory, so there may be a connection between the use of non-additive measures in quantum theory and interdependent complementary choices.[TODO: find citation that sub-additive is when choices are complementary] Both the Choquet and the Sugeno integral, for example, are defined over a monotonic set-functions, so there is a chance that these tools may turn out to be useful.
 We see this work as starting this line of research, which will take more effort to investigate.
@@
 A final remark before we start.
-Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects.
-Therefore, it is important to us that this work is a piece of a much larger puzzle, which connects cleanly and directly to  other areas, such as information geometry, functional analysis, quantum logic and so on.
+Too often, works in the foundations of physics focus on a particular aspect without concern for a broader connection to all relevant mathematical, physical, and philosophical aspects.
+Therefore, it is important to us that this work is a piece of a much larger puzzle, which connects cleanly and directly to other areas, such as information geometry, functional analysis, quantum logic, and so on.
 It is the hallmark of a true fundamental structure for physics that such connections can be easily made.
 Unfortunately, the nature of scientific publication requires us to publish a narrow slice of the work to fit the scope of a particular journal.
-Ironically, this means that the most important motivations for our approach are out of scope of this article, as we will concentrate only to the connections to measure theory.
+Ironically, this means that the most important motivations for our approach are out of scope of this article, as we will concentrate only on the connections to measure theory.
 
@@
 
-[This section has to make clear in what extra conditions on ensembles spaces Choquet theory can be applied.]
+[This section has to make clear in what extra conditions on ensemble spaces Choquet theory can be applied.]
 
@@
 That is, all statistical properties of $\ens$ can be recovered by the probability measure.
-This is what allows to think go $\ens$ as a probability distribution.
+This is what allows us to think of $\ens$ as a probability distribution.
 Note, however, that the Choquet theorem does not guarantee uniqueness.
 The probability measure is unique if and only if $E$ is a Choquet simplex, which is a generalization of the finite dimensional simplex.
-The characterization of the Choquet simplex is fairly technical, so we will simply define the Choque simplex by the property of having a single decomposition in terms of extreme points.
+The characterization of the Choquet simplex is fairly technical, so we will simply define the Choquet simplex by the property of having a single decomposition in terms of extreme points.
 
 % Limitation of applying Choquet theory to our case
-We should now look at the requirements of the theorem, to see whether we can always apply it to an ensemble space.
-First of all, and ensemble space $\Ens$ is a convex subset of a convex sapce $V$.
+We should now look at the requirements of the theorem to see whether we can always apply it to an ensemble space.
+First of all, an ensemble space $\Ens$ is a convex subset of a convex space $V$.
 Moreover, since every Hausdorff first-countable topological vector space is metrizable, $\Ens$ is metrizable.
@@
 Consider the space of probability measures over $\mathbb{R}$.
-Since $\mathbb{R}$ is not compact, the space of probability measure over $\mathbb{R}$ cannot be compact.\footnote{TODO: Check with mathematicians} Moreover, the set of extreme points is given by the Dirac measures (i.e.
+Since $\mathbb{R}$ is not compact, the space of probability measures over $\mathbb{R}$ cannot be compact.\footnote{TODO: Check with mathematicians} Moreover, the set of extreme points is given by the Dirac measures (i.e.
 probability one over a single point), which would have minus infinite entropy, and therefore cannot be part of the ensemble space.
-Therefore the theorem does not apply in general.
+Therefore, the theorem does not apply in general.
 
@@
 the Dirac measure) that are not, even conceptually, realizable.
-Many interesting cases fall into this restriction: the ensemble spaces of classical probability distribution over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
+Many interesting cases fall into this restriction: the ensemble spaces of classical probability distributions over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
 
 Note that ensembles over classical spaces will be represented by a unique probability measure, therefore the corresponding closure of the ensemble space is a Choquet simplex.
-However, quantum mixed states can be decomposed in multiple ways in terms of pure states, and therefore the closure of a quantum ensemble space is not a simplex.
+However, quantum mixed states can be decomposed in multiple ways in terms of pure states, and therefore, the closure of a quantum ensemble space is not a simplex.
 
@@
 Another way to represent quantum ensembles is through the use of quasiprobability distributions.
-Given a topological space $X$, a measure $\mu: \mathcal{B}(X) \to \mathbb{R}$ is affine if $\mu(X)=1$.
-All probability measures are affine measures, but the reverse is not true as affine measures can be negative on particular regions.
+Given a topological space $X$, a measure $\mu: \mathcal{B}(X) \to \mathbb{R}$ is affine if $\mu(X)=1$, where $\mathcal{B}(X)$ is the Borel $\sigma$-algebra.
+All probability measures are affine measures, but the reverse is not true, as affine measures can be negative on particular regions.
 The Wigner function, for example, can be negative over small regions of phase space.
 
-Given a space $X$, the space of all affine measure is a vector space, not just a convex subset as in the case of the space of probability measures.
+Given a space $X$, the space of all affine measures is a vector space, not just a convex subset as in the case of the space of probability measures.
 This means that, given an ensemble space, we may find a space $X$ such that the space of affine measures over $X$ is isomorphic to the embedding vector space $V$ as topological vector spaces.
@@
 Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
-Note that, given the uniform distribution, $1/6$ is exactly the probably for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
+Note that, given the uniform distribution, $1/6$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
 This gives the basic insight for our definitions.
@@
 The fraction is always well-defined because we can always write $\ens = 0 \ens[a] + 1 \ens$, therefore it must be zero or greater.
-This quantity tells us how much of ensemble $\ens$ can be constructed from $\ens[a]$.
+This quantity tells us how much of the ensemble $\ens$ can be constructed from $\ens[a]$.
 
@@
 Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.
-mixtures),\footnote{Since we work with convex sets within a topology, the hull is both the closed under the topology and convex combinations.
+mixtures),\footnote{Since we work with convex sets within a topology, the hull is both the closure under the topology and convex combinations.
 That is, $\hull(A)$ is the smallest closed and convex set that contains $A$} we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
@@
 For sub-additivity, note that the fraction of two components can at most sum during mixing.
-For the last, increasing and decreasing sequences strictly add or remove possible mixtures, therefore they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
+For the last, increasing and decreasing sequences strictly add or remove possible mixtures; therefore, they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
 The limit of these sequences must agree with the fraction capacity of the limit.
@@
 The fraction capacity, in fact, is not just defined on the extreme points, but it is defined on the whole space of ensembles.
-In the classical discrete case, for example, if $A \subset \Omega \subset \Ens$ is an event, then $\frcap_{\ens}(A)$ will return the probability of event $A$; if $P \subset \Ens$ represents all the Poisson distributions with integer coefficient, $frcap_{\ens}(P)$ will return how much of $\ens$ can be expressed is a mixture of those distributions.
+In the classical discrete case, for example, if $A \in \Omega \ni \Ens$ is an event, then $\frcap_{\ens}(A)$ will return the probability of event $A$; if $P \subset \Ens$ represents all the Poisson distributions with integer coefficient, $\frcap_{\ens}(P)$ will return how much of $\ens$ can be expressed as a mixture of those distributions.
 Similarly, the probability measures in Choquet theory are, in general, defined on the whole ensemble space, and then one finds those particular ones that have support on the extreme points.
@@
 This means that each element of the ensemble space is represented by a unique probability measure over the extreme points.
-In that case, the fraction capacity extended on the extreme points will also be additive.
+In that case, the fraction capacity, extended to the extreme points, will also be additive.
 In the non-classical case, the closure of the ensemble space will not be a simplex, and there will be multiple representations, which will lead to a non-additive measure.
 Suppose, in fact, that $\mu$ and $\nu$ are two distinct measures that represent the same ensemble $\ens$.
-Then there will be a set $A$ such such that $\mu(A) \neq \nu(A)$.
+Then there will be a set $A$ such that $\mu(A) \neq \nu(A)$.
 Suppose, without loss of generality, that $\mu(A) > \nu(A)$ We will also have $\mu(A^{\complement}) = 1- \mu(A) < 1 - \nu(A) = \nu(A^{\complement})$.
@@
 Since entropy characterizes the variability of the preparations of the ensemble, greater variability means the ensemble is spread over more cases.
-Therefore the count of distinguishable states must be a monotonic function of the entropy.
+Therefore, the count of distinguishable states must be a monotonic function of the entropy.
 It has been shown[cite Hall] under very general conditions that the exponential of the entropy is the only function to have the appropriate features.
@@
 The first property readily comes from the range of the exponential.
-For the second, note that $\hull$, exponential and $\sup$ are all monotone.
+For the second, note that $\hull$, exponential, and $\sup$ are all monotone.
 For sub-additivity, the exponential of the entropy itself is sub-additive, and it is additive over orthogonal ensembles.
@@
 
-Again we note that the non-additive measure we recover is very well-behaved, and connects directly with important quantities.
+Again, we note that the non-additive measure we recover is very well-behaved and connects directly with important quantities.
 When Choquet theory applies, state capacity can be extended to the extreme points.
 In the classical case, disjoint sets of extreme points will be orthogonal, meaning that the state capacity there restricted is an additive measure.
-In the discrete case, every point is associated to zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
-In the continuous case, gives a subset of phase space $A$, the maximum entropy is reached by the uniform distribution $\rho_A$, which means we recover $\log_2 \capacity(A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
+In the discrete case, every point is associated with zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
+In the continuous case, given a subset of phase space $A$, the maximum entropy is reached by the uniform distribution $\rho_A$, which means we recover $\log_2 \capacity(A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
 Similarly, in quantum mechanics, given an orthogonal subspace $A$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(A) = 2^{\dim(A)}$.
@@
 In classical mechanics, all states are perfectly distinguishable.
-In quantum mechanics, only orthogonal state are distinguishable, and therefore the state capacity recovers the number of orthogonal states.
+In quantum mechanics, only orthogonal states are distinguishable, and therefore the state capacity recovers the number of orthogonal states.
 

### 2025-06-27 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
For submission to Nature Physics
Fix bibliography capitalization and typo

@@
 \section{Introduction}
+\label{sec:intro}
 
@@
 One of the experimental challenges in quantum mechanics is producing states that are ``quantum enough'' to exhibit quantum properties.
-What we want to show that all these problems can ultimately be understood as reducing the entropy of the initial state.
+What we want to show is that all these problems can ultimately be understood as reducing the entropy of the initial state.
 
@@
 In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state~\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive.
-In section 4, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite-dimensional case, one could have infinite-dimensional subspaces that remain orthogonal while increasing the entropy of all its elements.
+In section~\ref{sec:high_entropy_limit}, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite-dimensional case, one could have infinite-dimensional subspaces that remain orthogonal while increasing the entropy of all its elements.
 This corresponds, for example, to the case where there are conserved quantities.
@@
 
+\section*{Acknowledgements}
+This paper is part of the ongoing Assumptions of Physics project~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+
+% TODO: include funding information
+\paragraph{Funding information}
+ML acknowledges funding via the FWF-funded QuantERA project QuSiED with project number I 6008-N.
+
 \bibliographystyle{plain}

### 2025-07-01 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: worked on the pseudo-probability section

--- 2025-06-04 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-Another way to represent quantum ensembles is through the use of quasiprobability distributions.
-Given a topological space $X$, a measure $\mu$ is affine if $\mu(X)=1$.
-All probability measures are affine measures, but the reverse is not true as affine measures can be negative on particular regions.
-The Wigner function, for example, can be negative over small regions of phase space.
+%Goal of this section: quasi-probability representations are allowed by the embedding in a topological vector space.
+Choice of representations are guided by what observables are interesting for a specific problem.
+
+%Examples: Wigner, Husimi, Glauber P, SIC-POVM (space not made of spectra of operators), Kirkwood-Dirac (two-time correlation function - observables not necessarily taken at the same time )
+
+%Things to say: some people think about it as weakining axioms (cite papers on negative probability).
+It's more than that.
+Points describe events that are not necessarily mutually exclusive.
+TODO: example of two Wigner function/Husimi Q would be better that have no overlap but they are not orthogonal and/or two functions that have overlap but are orthogonal.
+
+%Thing to say: all these representations are useful because they preserve the linear structure, meaning that convex combinations of ensembles are convex combinations of quasi-probability; expectation values are linear.
+Check, but it should be true.
+
+% 1 describe quasiprobability approaches; go through examples; point out that the quasi-probability can be negative; that one can recover expectations;
+
+\textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions.
+While the Wigner function is probably the most widely known, there are in fact several different approach.
+The Wigner function represent quantum states, both pure and mixed, as a density function over position and momentum.
+This function integrates to one over the full space, like a probability density, but can have small negative regions, unlike a probability density.
+Recall that an affine combination is a linear combination where all coefficients sum to one but are not necessarily non-negative.
+Therefore, in general, quasi-probability distributions can be understood as affine combinations.
+
+For the Wigner function, the expectation of polynomials of position and momentum over phase space recovers expectations of polynomials of position and momentum in symmetric order (?).
+(Definitions and features?) In some problems, especially in quantum optics, normal ordering and anti-normal ordering are more useful.
+The Glauber P representation and the Husimi Q representation work like the Wigner function except that the expectations of polynomials recover the respective order.
+In other circumstances, quasi-probability distributions over observables different from position and momentum are more useful.
+For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times.
+Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
+For example, the SIC-POVM represents density matrices as affine combinations of pure states that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
+
+It should be clear that there are many ways to represent quantum states via quasi-probability distributions, and that the choice of representation is a matter of convenience for the problem at hand.
+All representations, though, have the following features in common.
+First, we have a map $W : \Ens \to Q(Z,\mathbb{R})$ that given each mixed state $\rho \in \Ens$ returns the corresponding quasi-probability function $w(z) \in Q(Z,\mathbb{R})$, where $Q(Z,\mathbb{R})$ is the function space used by the specific representation.
+The representation preserves convex combinations: $W(\sum_i p_i \rho_i) = \sum_i p_i W(\rho_i)$.
+That is, a statistical mixture of quasi-probability distributions represents the statistical mixture of the corresponding states.
+Since the map is linear, and expectation values are linear, one can express expectation values as integrals or sums over the quasi-probability $w(z)$.
+
+% 2 Note that it's not about negative probability; disjoint not mutually exclusive
+
+\textbf{Not about negative probability.} It is very common to conceive the main difference between classical probability and these quasi-probabilities as the inclusion of negative probability.
+Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
+This is a mischaracterization.
+For example, the Husimi Q representation provides us with a non-negative quasi-probability function.
+That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a probability distribution.
+
+The crucial difference is that in classical probability all points of the sample space represent mutually exclusive events.
+That is, they can be prepared and observed independently from the others.
+Therefore, any two events that are disjoint are also mutually exclusive.
+This is not the case for quasi-probability in quantum mechanics.
+Husimi Q represents a state as decomposed over coherent states, which cannot represent mutually exclusive as they are not orthogonal.
+Similarly, we can imagine to embed the Bloch ball into a tetrahedron, which means every point in the Bloch ball could be represented as a convex combination of extreme points, meaning that they would be represented by a non-negative quasi-probability distribution over four elements.
+However, none of the four elements represent a valid state.
+
+% 3 These representations are possible because of the vector space structure of the ensemble space, and therefore something similar will be possible in ANY physical theory
 
@@
 
+% 4 pros and cons - can choose a repretation for the problem at hand; no fundamental representation; boundaries ore not well defined (i.e. which functions represnt valid states)
+
 This alternative way of looking at measure theoretic representations, then, is consistent with the previous one.
@@
 To develop a general theory for physics, however, this is not useful as each representation is not general enough.
-
-TODO: disjoint supports of Wigner functions do not give orthogonal subspaces (CHECK)
 

### 2025-07-02 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Cleanup first pages

--- 2025-06-10 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-\title{A non-additive generalization of probability theory \\for quantum mechanics and beyond}
+\title{A non-additive generalization of probability theory\\for quantum mechanics and beyond}
 \author{Gabriele Carcassi}
@@
 The purpose of this paper is to show the core ideas and present questions that may be developed on the mathematical, physical and philosophical side.
+
 \end{abstract}
@@
 
-\section{TOC}
+\tableofcontents
 
-Overall goal: to create a generalization of probability theory and measure theory for general physical systems.
-
-General question: how do we represent physical states? Different types of representations: talk about examples, ....
-
--- General problem
-
-For this work: assume we have a locally convex topological vector space V. E is a bounded convex subset of V. Equipped with an entropy. Topology generated by entropic balls?
-
-Separateness and orthogonality
-
-Statistical quantities
-
-Show that they reduce to standard things in classical mechanics and quantum mechanics.
-
--- Connection to Choquet theory
-
-Choquet theory - in the case that the closure of the space is compact -> representation through measures. May not be unique
-
-Choquet simplex. Uniqueness of decomposition and separateness/orthogonality.
-
--- Connection to signed probability
-
-TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
-
-Characterize signed measures as an affine representation of the ensemble space (i.e. bounded convex subset of a locally convex topological vector space).
-
-Show that the probability is signed if and only if multiple decomposition, non-additive, no Choquet simplex
-
-Signed probability is unique, but does not characterize the convex set (ensemble space).
+%\section{TOC}
+%
+%Overall goal: to create a generalization of probability theory and measure theory for general physical systems.
+%
+%General question: how do we represent physical states? Different types of representations: talk about examples, ....
+%
+%-- General problem
+%
+%For this work: assume we have a locally convex topological vector space V. E is a bounded convex subset of V. Equipped with an entropy. Topology generated by entropic balls?
+%
+%Separateness and orthogonality
+%
+%Statistical quantities
+%
+%Show that they reduce to standard things in classical mechanics and quantum mechanics.
+%
+%-- Connection to Choquet theory 
+%
+%Choquet theory - in the case that the closure of the space is compact -> representation through measures. May not be unique
+%
+%Choquet simplex. Uniqueness of decomposition and separateness/orthogonality.
+%
+%-- Connection to signed probability
+%
+%TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
+%
+%Characterize signed measures as an affine representation of the ensemble space (i.e. bounded convex subset of a locally convex topological vector space).
+%
+%Show that the probability is signed if and only if multiple decomposition, non-additive, no Choquet simplex
+%
+%Signed probability is unique, but does not characterize the convex set (ensemble space).
 Note: the other measures are defined on the extreme points.
@@
 The non-additive measure does not characterize the extreme points, so it doesn't really characterize the space.
-
--- Our non-additive measure approach
-
-State capacity
-
-Fraction capacity. Show uniqueness. Also show uniqueness on extreme points.
-
-Fraction capacity as supremum of all possible measures
-
-Additivity of fraction capacity (and state capacity).
-
-Single decomposition and classical spaces and measurement contexts
-
-Show that they reduce to standard things in classical mechanics and quantum mechanics.
-
--- Connection to fuzzy logic?
-
-Literature search.
-
-Possible connection to
-* Choquet integral
-* Sugeno integral
-
-? Space of fuzzy measure like we have a space of additive measures?
+%
+%-- Our non-additive measure approach
+%
+%State capacity
+%
+%Fraction capacity. Show uniqueness. Also show uniqueness on extreme points.
+%
+%Fraction capacity as supremum of all possible measures
+%
+%Additivity of fraction capacity (and state capacity).
+%
+%Single decomposition and classical spaces and measurement contexts
+%
+%Show that they reduce to standard things in classical mechanics and quantum mechanics.
+%
+%-- Connection to fuzzy logic?
+%
+%Literature search.
+%
+%Possible connection to
+%* Choquet integral
+%* Sugeno integral
+%
+%? Space of fuzzy measure like we have a space of additive measures?
 

### 2025-07-02 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Add Kolmogorov axioms

@@
 \usepackage{tcolorbox}
+\usepackage{scrextend}	% for environment ``labeling''
 
@@
 Another way to represent quantum ensembles is through the use of quasiprobability distributions.
-Given a topological space $X$, a measure $\mu: \mathcal{B}(X) \to \mathbb{R}$ is affine if $\mu(X)=1$, where $\mathcal{B}(X)$ is the Borel $\sigma$-algebra.
-All probability measures are affine measures, but the reverse is not true, as affine measures can be negative on particular regions.
-The Wigner function, for example, can be negative over small regions of phase space.
+These distributions (or measures) relax some of Kolmogorov's probability axioms\footnote{TODO: These should be stated earlier.}.
+
+A measure $\mu: \Omega_X \to \mathbb{R}$ over a sample space $X$ with $\sigma$-algebra $\Omega_X$ is called a probability measure if it satisfies the following three axioms
+\begin{labeling}{Non-negativity:}
+	\item[Non-negativity:] $\forall x \in \Omega_X: \mu(x) \ge 0$
+	\item[Normalization:] $\mu(X) = 1$
+	\item[$\sigma$-additivity:] $\mu ( \bigcup_i x_i ) = \sum_i \mu ( x_i )$ for any countable sequence of disjoint sets $x_i \in \Omega_X$.
+\end{labeling}
+
+For example, affine measures do not necessarily satisfy non-negativity. The Wigner function is the probability density of such a measure and can be negative over small regions of phase space.
 

### 2025-07-02 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex

@@
 
-Another way to represent quantum ensembles is through the use of quasiprobability distributions.
-These distributions (or measures) relax some of Kolmogorov's probability axioms\footnote{TODO: These should be stated earlier.}.
+%Goal of this section: quasi-probability representations are allowed by the embedding in a topological vector space.
+Choice of representations are guided by what observables are interesting for a specific problem.
 
-A measure $\mu: \Omega_X \to \mathbb{R}$ over a sample space $X$ with $\sigma$-algebra $\Omega_X$ is called a probability measure if it satisfies the following three axioms
+%Examples: Wigner, Husimi, Glauber P, SIC-POVM (space not made of spectra of operators), Kirkwood-Dirac (two-time correlation function - observables not necessarily taken at the same time )
+
+%Things to say: some people think about it as weakining axioms (cite papers on negative probability).
+It's more than that.
+Points describe events that are not necessarily mutually exclusive.
+TODO: example of two Wigner function/Husimi Q would be better that have no overlap but they are not orthogonal and/or two functions that have overlap but are orthogonal.
+
+A measure $\mu: \Omega_X \to \mathbb{R}$ over a sample space $X$ with $\sigma$-algebra $\Omega_X$ is called a probability measure if it satisfies the following three axioms\footnote{TODO: These should be stated earlier.}
 \begin{labeling}{Non-negativity:}
@@
 \end{labeling}
+
+%Thing to say: all these representations are useful because they preserve the linear structure, meaning that convex combinations of ensembles are convex combinations of quasi-probability; expectation values are linear.
+Check, but it should be true.
+
+% 1 describe quasiprobability approaches; go through examples; point out that the quasi-probability can be negative; that one can recover expectations;
+
+\textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions.
+While the Wigner function is probably the most widely known, there are in fact several different approach.
+The Wigner function represent quantum states, both pure and mixed, as a density function over position and momentum.
+This function integrates to one over the full space, like a probability density, but can have small negative regions, unlike a probability density.
+Recall that an affine combination is a linear combination where all coefficients sum to one but are not necessarily non-negative.
+Therefore, in general, quasi-probability distributions can be understood as affine combinations.
+
+For the Wigner function, the expectation of polynomials of position and momentum over phase space recovers expectations of polynomials of position and momentum in symmetric order (?).
+(Definitions and features?) In some problems, especially in quantum optics, normal ordering and anti-normal ordering are more useful.
+The Glauber P representation and the Husimi Q representation work like the Wigner function except that the expectations of polynomials recover the respective order.
+In other circumstances, quasi-probability distributions over observables different from position and momentum are more useful.
+For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times.
+Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
+For example, the SIC-POVM represents density matrices as affine combinations of pure states that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
+
+It should be clear that there are many ways to represent quantum states via quasi-probability distributions, and that the choice of representation is a matter of convenience for the problem at hand.
+All representations, though, have the following features in common.
+First, we have a map $W : \Ens \to Q(Z,\mathbb{R})$ that given each mixed state $\rho \in \Ens$ returns the corresponding quasi-probability function $w(z) \in Q(Z,\mathbb{R})$, where $Q(Z,\mathbb{R})$ is the function space used by the specific representation.
+The representation preserves convex combinations: $W(\sum_i p_i \rho_i) = \sum_i p_i W(\rho_i)$.
+That is, a statistical mixture of quasi-probability distributions represents the statistical mixture of the corresponding states.
+Since the map is linear, and expectation values are linear, one can express expectation values as integrals or sums over the quasi-probability $w(z)$.
+
+% 2 Note that it's not about negative probability; disjoint not mutually exclusive
+
+\textbf{Not about negative probability.} It is very common to conceive the main difference between classical probability and these quasi-probabilities as the inclusion of negative probability.
+Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
+This is a mischaracterization.
+For example, the Husimi Q representation provides us with a non-negative quasi-probability function.
+That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a probability distribution.
+
+The crucial difference is that in classical probability all points of the sample space represent mutually exclusive events.
+That is, they can be prepared and observed independently from the others.
+Therefore, any two events that are disjoint are also mutually exclusive.
+This is not the case for quasi-probability in quantum mechanics.
+Husimi Q represents a state as decomposed over coherent states, which cannot represent mutually exclusive as they are not orthogonal.
+Similarly, we can imagine to embed the Bloch ball into a tetrahedron, which means every point in the Bloch ball could be represented as a convex combination of extreme points, meaning that they would be represented by a non-negative quasi-probability distribution over four elements.
+However, none of the four elements represent a valid state.
+
+% 3 These representations are possible because of the vector space structure of the ensemble space, and therefore something similar will be possible in ANY physical theory
 

### 2025-07-02 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: more on the section about pseudo-probability

@@
 \usepackage{tcolorbox}
-\usepackage{scrextend}	% for environment ``labeling''
 
@@
 
-\title{A non-additive generalization of probability theory\\for quantum mechanics and beyond}
+\title{A non-additive generalization of probability theory \\for quantum mechanics and beyond}
 \author{Gabriele Carcassi}
@@
 \begin{abstract}
-	We present a physically motivated generalization of probability theory that is suitable for classical mechanics, quantum mechanics, and any future physical theory that allows a statistical description.
-The goal is to put the use of classical and quantum probability in a broader context and show how the current mathematical structures are likely not suitable to solve the open problems in the foundations of physics.
-For the more mathematically inclined, we will point to areas where new math or generalization of established math is needed.
-For the more philosophically inclined, we will point to areas where further conceptual work is needed.
+	We present a physically motivated generalization of probability theory that is suitable for classical mechanics, quantum mechanics and any future physical theory that allows a statistical description.
+The goal is to put the use of classical and quantum probability in a broader context, and show how the current mathematical structures are likely not suitable to solve the open problems in the foundations of physics.
+For the more mathematically inclined, we will point to areas where new math or generalization of established math are needed.
+For the more philosophically incline, we will point to areas where further conceptual work is needed.
 
@@
 The purpose of this paper is to show the core ideas and present questions that may be developed on the mathematical, physical and philosophical side.
-
 \end{abstract}
@@
 
-\tableofcontents
+\section{TOC}
 
-%\section{TOC}
-%
-%Overall goal: to create a generalization of probability theory and measure theory for general physical systems.
-%
-%General question: how do we represent physical states? Different types of representations: talk about examples, ....
-%
-%-- General problem
-%
-%For this work: assume we have a locally convex topological vector space V. E is a bounded convex subset of V. Equipped with an entropy. Topology generated by entropic balls?
-%
-%Separateness and orthogonality
-%
-%Statistical quantities
-%
-%Show that they reduce to standard things in classical mechanics and quantum mechanics.
-%
-%-- Connection to Choquet theory 
-%
-%Choquet theory - in the case that the closure of the space is compact -> representation through measures. May not be unique
-%
-%Choquet simplex. Uniqueness of decomposition and separateness/orthogonality.
-%
-%-- Connection to signed probability
-%
-%TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
-%
-%Characterize signed measures as an affine representation of the ensemble space (i.e. bounded convex subset of a locally convex topological vector space).
-%
-%Show that the probability is signed if and only if multiple decomposition, non-additive, no Choquet simplex
-%
-%Signed probability is unique, but does not characterize the convex set (ensemble space).
+Overall goal: to create a generalization of probability theory and measure theory for general physical systems.
+
+General question: how do we represent physical states? Different types of representations: talk about examples, ....
+
+-- General problem
+
+For this work: assume we have a locally convex topological vector space V. E is a bounded convex subset of V. Equipped with an entropy. Topology generated by entropic balls?
+
+Separateness and orthogonality
+
+Statistical quantities
+
+Show that they reduce to standard things in classical mechanics and quantum mechanics.
+
+-- Connection to Choquet theory
+
+Choquet theory - in the case that the closure of the space is compact -> representation through measures. May not be unique
+
+Choquet simplex. Uniqueness of decomposition and separateness/orthogonality.
+
+-- Connection to signed probability
+
+TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
+
+Characterize signed measures as an affine representation of the ensemble space (i.e. bounded convex subset of a locally convex topological vector space).
+
+Show that the probability is signed if and only if multiple decomposition, non-additive, no Choquet simplex
+
+Signed probability is unique, but does not characterize the convex set (ensemble space).
 Note: the other measures are defined on the extreme points.
@@
 The non-additive measure does not characterize the extreme points, so it doesn't really characterize the space.
-%
-%-- Our non-additive measure approach
-%
-%State capacity
-%
-%Fraction capacity. Show uniqueness. Also show uniqueness on extreme points.
-%
-%Fraction capacity as supremum of all possible measures
-%
-%Additivity of fraction capacity (and state capacity).
-%
-%Single decomposition and classical spaces and measurement contexts
-%
-%Show that they reduce to standard things in classical mechanics and quantum mechanics.
-%
-%-- Connection to fuzzy logic?
-%
-%Literature search.
-%
-%Possible connection to
-%* Choquet integral
-%* Sugeno integral
-%
-%? Space of fuzzy measure like we have a space of additive measures?
+
+-- Our non-additive measure approach
+
+State capacity
+
+Fraction capacity. Show uniqueness. Also show uniqueness on extreme points.
+
+Fraction capacity as supremum of all possible measures
+
+Additivity of fraction capacity (and state capacity).
+
+Single decomposition and classical spaces and measurement contexts
+
+Show that they reduce to standard things in classical mechanics and quantum mechanics.
+
+-- Connection to fuzzy logic?
+
+Literature search.
+
+Possible connection to
+* Choquet integral
+* Sugeno integral
+
+? Space of fuzzy measure like we have a space of additive measures?
 
@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used, given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
 
 We reformulate the problem in the following way: given the space of all possible statistical ensembles allowed by a statistical theory, what measure theoretic tools can we use to represent each ensemble and the geometric properties of the overall space?
-Why exactly do classical probability measures fail to capture completely quantum states?
+Why exactly classical probability measures fail to capture completely quantum states?
 What suitable generalizations are possible, and what mathematical tools would need to be developed?
@@
 These mathematical properties are physically grounded in the ability to experimentally define ensemble (leading to a topological space), to define statistical mixtures (leading to a convex space), to quantify the variability of the elements of the ensembles (leading to an entropy function) and to having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
-The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurement (as in the dimension of orthogonal subspaces of an operator)?
+The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the dimension of orthogonal subspaces of an operator)?
 What measure theoretic tool can represent each ensemble?
@@
 We will first review some results from Choquet theory, which is, to our knowledge, the most general approach to represent elements of a convex set using additive probability measures.
-The key insight is that, at least in the compact case, such a representation is always possible, but it is not unique.
+The key insight is that, at least in the compact case, such representation is always possible, but it is not unique.
 Uniqueness requires the space to be a Choquet simplex, which is a generalization of the standard finite-dimensional simplex and is a particularly shaped subset of a vector space.
@@
 That is, there is no unique or privileged choice for this representation.
-Moreover, since the map is at the level of the full vector space, not just the subset with valid ensembles, there is no general way to characterize which subset of signed probability measures corresponds to ensembles.
+Moreover, since the map is at the level of the full vector space, not just the subset with valid ensembles, there is no general way to characterize which subset of signed probability measures correspond to ensembles.
 Therefore, while signed measures are a powerful tool for calculation in specific problems, they cannot give us a general principled representation of the physical theory.
@@
 The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
-From the entropy, we can define the state capacity, which generalizes the notion of the count of states.
+From the entropy, we can define the state capacity, which generalizes the notion of count of states.
 For each ensemble, we can define the fraction capacity, which generalizes the notion of probability.
 Both are non-negative sub-additive monotonic set functions, with the second being unit bounded.
-These measures are always unique and, since they are defined on the ensemble space itself, always provide a physically meaningful representation.
-Moreover, in the compact case, these measures can be restricted to the extreme points while retaining uniqueness.
+These measures are always unique and, since they are defined on the ensemble space itself, always have provide a physically meaningful representation.
+Moreover, in the compact case, these measure can be restricted on the extreme points while retaining uniqueness.
 (TODO: prove)
@@
 The fraction capacity is the supremum of all possible measures for the same ensemble.
-The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure and is therefore additive.
+The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure, and is therefore additive.
 The state capacity, instead, corresponds to the minimum of the entropy of all possible representations.
@@
 Having shown that a general approach based on non-additive measures is possible, we lay down a series of questions to explore in the future.
-The biggest one is whether it is possible to generalize notions such as integration, derivation, and expectation to this non-additive case.
-It should be noted that non-additive measures are used in the context of decision theory, so there may be a connection between the use of non-additive measures in quantum theory and interdependent complementary choices.[TODO: find citation that sub-additive is when choices are complementary] Both the Choquet and the Sugeno integral, for example, are defined over a monotonic set-functions, so there is a chance that these tools may turn out to be useful.
+The biggest one is whether it is possible to generalize notions such as integration, derivation and expectation to this non-additive case.
+It should be noted that non-additive measures are used in the context of decision theory, so there may be a connection between the use of non-additive measures in quantum theory and interdependent complimentary choices.[TODO: find citation that sub-additive is when choices are complementary] Both the Choquet and the Sugeno integral, for example, are defined over a monotonic set-functions, so there is a chance that these tools may turn out to be useful.
 We see this work as starting this line of research, which will take more effort to investigate.
@@
 A final remark before we start.
-Too often, works in the foundations of physics focus on a particular aspect without concern for a broader connection to all relevant mathematical, physical, and philosophical aspects.
-Therefore, it is important to us that this work is a piece of a much larger puzzle, which connects cleanly and directly to other areas, such as information geometry, functional analysis, quantum logic, and so on.
+Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects.
+Therefore, it is important to us that this work is a piece of a much larger puzzle, which connects cleanly and directly to  other areas, such as information geometry, functional analysis, quantum logic and so on.
 It is the hallmark of a true fundamental structure for physics that such connections can be easily made.
 Unfortunately, the nature of scientific publication requires us to publish a narrow slice of the work to fit the scope of a particular journal.
-Ironically, this means that the most important motivations for our approach are out of scope of this article, as we will concentrate only on the connections to measure theory.
+Ironically, this means that the most important motivations for our approach are out of scope of this article, as we will concentrate only to the connections to measure theory.
 
@@
 
-Additionally, each ensemble space $\Ens$ is equipped with a continuous entropy function $S : \Ens \to \mathbb{R}$. The entropy of a statistical mixture satisfies the following bounds
+Additional, each ensemble space $\Ens$ is equipped with a continuous entropy function $S : \Ens \to \mathbb{R}$. The entropy of a statistical mixture satisfies the following bounds
 \begin{equation}
@@
 Intuitively, the variability cannot decrease during mixing, and will have an upper bound when mixing two ensembles that are non-overlapping, or orthogonal.
-In fact, we define two ensembles $\ens[a] \ortho \ens[b]$ to be orthogonal when their mixture gives a maximal entropy increase.
+In fact, we define two ensembles $\ens[a] \ortho \ens[b]$ to be orthogonal when they mixture gives a maximal entropy increase.
 This does recover the notion of orthogonality in the corresponding classical and quantum vector spaces of statistical ensembles.
@@
 
-[This section has to make clear in what extra conditions on ensemble spaces Choquet theory can be applied.]
+[This section has to make clear in what extra conditions on ensembles spaces Choquet theory can be applied.]
 
@@
 That is, all statistical properties of $\ens$ can be recovered by the probability measure.
-This is what allows us to think of $\ens$ as a probability distribution.
+This is what allows to think go $\ens$ as a probability distribution.
 Note, however, that the Choquet theorem does not guarantee uniqueness.
 The probability measure is unique if and only if $E$ is a Choquet simplex, which is a generalization of the finite dimensional simplex.
-The characterization of the Choquet simplex is fairly technical, so we will simply define the Choquet simplex by the property of having a single decomposition in terms of extreme points.
+The characterization of the Choquet simplex is fairly technical, so we will simply define the Choque simplex by the property of having a single decomposition in terms of extreme points.
 
 % Limitation of applying Choquet theory to our case
-We should now look at the requirements of the theorem to see whether we can always apply it to an ensemble space.
-First of all, an ensemble space $\Ens$ is a convex subset of a convex space $V$.
+We should now look at the requirements of the theorem, to see whether we can always apply it to an ensemble space.
+First of all, and ensemble space $\Ens$ is a convex subset of a convex sapce $V$.
 Moreover, since every Hausdorff first-countable topological vector space is metrizable, $\Ens$ is metrizable.
@@
 Consider the space of probability measures over $\mathbb{R}$.
-Since $\mathbb{R}$ is not compact, the space of probability measures over $\mathbb{R}$ cannot be compact.\footnote{TODO: Check with mathematicians} Moreover, the set of extreme points is given by the Dirac measures (i.e.
+Since $\mathbb{R}$ is not compact, the space of probability measure over $\mathbb{R}$ cannot be compact.\footnote{TODO: Check with mathematicians} Moreover, the set of extreme points is given by the Dirac measures (i.e.
 probability one over a single point), which would have minus infinite entropy, and therefore cannot be part of the ensemble space.
-Therefore, the theorem does not apply in general.
+Therefore the theorem does not apply in general.
 
 % Define "Choquet space" as those spaces where we can apply Choquet theory
-We can apply Choquet's theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
+We can apply Choquet theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
 TODO: finalize the proof} and metrizable.
-In this case, we are guaranteed that all ensembles allow a representation through a probability on points that may correspond to idealized ensembles (i.e.
+In this case, we are guaranteed that all ensembles allows a representation through a probability on points that may corresponds idealized ensembles (i.e.
 the Dirac measure) that are not, even conceptually, realizable.
-Many interesting cases fall into this restriction: the ensemble spaces of classical probability distributions over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
+Many interesting cases fall into this restriction: the ensemble spaces of classical probability distribution over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
 
 Note that ensembles over classical spaces will be represented by a unique probability measure, therefore the corresponding closure of the ensemble space is a Choquet simplex.
-However, quantum mixed states can be decomposed in multiple ways in terms of pure states, and therefore, the closure of a quantum ensemble space is not a simplex.
+However, quantum mixed states can be decomposed in multiple ways in terms of pure states, and therefore the closure of a quantum ensemble space is not a simplex.
 
@@
 
-Chouqet theory tells us, in the realm where it can be applied, that classical probability fails in quantum mechanics precisely because the ensemble space of a quantum system allows multiple decompositions.
+Chouqet theory tells, in the realm where it can be applied, that classical probability fails in quantum mechanics precisely because the ensemble space of a quantum system allows multiple decomposition.
 
@@
 TODO: example of two Wigner function/Husimi Q would be better that have no overlap but they are not orthogonal and/or two functions that have overlap but are orthogonal.
-
-A measure $\mu: \Omega_X \to \mathbb{R}$ over a sample space $X$ with $\sigma$-algebra $\Omega_X$ is called a probability measure if it satisfies the following three axioms\footnote{TODO: These should be stated earlier.}
-\begin{labeling}{Non-negativity:}
-	\item[Non-negativity:] $\forall x \in \Omega_X: \mu(x) \ge 0$
-	\item[Normalization:] $\mu(X) = 1$
-	\item[$\sigma$-additivity:] $\mu ( \bigcup_i x_i ) = \sum_i \mu ( x_i )$ for any countable sequence of disjoint sets $x_i \in \Omega_X$.
-\end{labeling}
 
@@
 For example, the Husimi Q representation provides us with a non-negative quasi-probability function.
-That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a probability distribution.
+That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.
+Similarly, since we can embed the Bloch ball inside a tetrahedron, we can represent all states of a two-state quantum system by a convex combination of the vertices, leading to a non-negative quasi-probability distribution over four elements.
 
-The crucial difference is that in classical probability all points of the sample space represent mutually exclusive events.
-That is, they can be prepared and observed independently from the others.
-Therefore, any two events that are disjoint are also mutually exclusive.
-This is not the case for quasi-probability in quantum mechanics.
+Conversely, we can use signed measure to represent classical distributions.
+Suppose we have a two state classical discrete space, which means the space of all possible ensembles is spanned by $[p, (1-p)]$.
+Alternatively, we can write it as $p[1,0] + (1-p)[0,1]$ which highlights that we are writing each possible probability distribution in terms of the two possible cases.
+But we could also express each probability distribution in terms of other two distributions, as in $\left(2p - \frac{1}{2}\right)\left[\frac{3}{4}, \frac{1}{4}\right] + \left(\frac{3}{2} - 2p\right)\left[\frac{1}{4}, \frac{3}{4}\right]$.
+The coefficients now span between $\left[-\frac{1}{2}, \frac{3}{2}\right]$, can be negative, but still sum to one.
+That is, in terms of this new affine basis, we need to use a signed probability.
+
+Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measure, the difference between classical probability and quantum quasi-probability does not lie in the use of negative probability.
+Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events.
+That is, they can be prepared and observed independently from the others; we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasi-probability in quantum mechanics.
 Husimi Q represents a state as decomposed over coherent states, which cannot represent mutually exclusive as they are not orthogonal.
-Similarly, we can imagine to embed the Bloch ball into a tetrahedron, which means every point in the Bloch ball could be represented as a convex combination of extreme points, meaning that they would be represented by a non-negative quasi-probability distribution over four elements.
-However, none of the four elements represent a valid state.
+The inner product between these states will provide a constrain.
+In the non-negative probability representation of the Bloch sphere, none of the four corners represent valid states, as they lie outside the Bloch ball.
+So we cannot assign probability to a vertex without assigning some probability to another vertex.
+
+While negative probability is a red herring, we want to understand two things.
+When can we use classical probability?
+Are quasi-probability representation specific to quantum mechanics, or are they more general?
+Fortunately, the ensemble space framework allows us to answer these questions directly.
 
@@
 
-For example, affine measures do not necessarily satisfy non-negativity. The Wigner function is the probability density of such a measure and can be negative over small regions of phase space.
+\textbf{Ensemble space embedding and signed measures.} We have already seen that we can represent ensembles using classical probability if and only the ensemble space is a simplex.
+Since any probability measure is possible, we can assign probability to two disjoint areas independently.
+This answers the first question.
 
-Given a space $X$, the space of all affine measures is a vector space, not just a convex subset as in the case of the space of probability measures.
-This means that, given an ensemble space, we may find a space $X$ such that the space of affine measures over $X$ is isomorphic to the embedding vector space $V$ as topological vector spaces.
-In this case, any element $v\in V$ can be represented uniquely by an affine measure $\mu_{v}$, regardless of whether they are actual ensembles or not.\footnote{The question of whether such an $X$ always exists is left open.
-We suspect a link exists between the statistical quantities $F_i$ that fully identify each ensemble and the space $X$.} Because the statistical quantities $F_i$ are affine functions, we will have $F_i(v) = \int_X F_i d\mu_{v}$.
+For the second question, let $Z$ be the space over which we are defining our quasi-probability distributions.
+Note that, in general, $Z$ is not the space of states.
+Wigner functions, for example, are defined over phase space, which does not represent quantum states.
+Let $M^{\pm}(Z)$ the space of all bounded signed measures over $Z$.
+Since the linear combination of two bounded signed measure is a bounded signed measure, $M^{\pm}(Z)$ is a vector space.
+Now let $M_1^{\pm}(Z)$ be the space of signed probability measures.
+While it is not a subspace of $M(Z)$, it does not include the zero measure, it is an affine subspace therefore, up to the choice of an arbitrary origin, it is a vector space.
 
-In general, since the ensemble space is bounded along every direction, not all affine measures will correspond to an ensemble space.
-Whether all probability measures correspond to a valid ensemble, or whether affine measures are even used, corresponds to relationships between $\Ens$ and the space of probability measures $M_p(X)$.
-If $\Ens \subset M_p(X)$, then all states are represented by a probability measure.
-However, this is not a ``probability'' in the standard sense of the word.
-For example, the Husimi Q representation is always non-negative at every point, but two points do not represent mutually exclusive events as they are not orthogonal (i.e.
-their inner product is non-zero and they do not maximize the entropy increase during mixture).
-Mathematically, the issue is that the points $X$ of the Husimi Q representation are not the extreme points of the closure ensemble space $\overline{\Ens}$.
-In general, $\Ens$ and $M_p(X)$ will overlap, but neither will contain the other.
-For example, in the Wigner representation, the only pure states (i.e.
-the extreme points of $\Ens$) represented by a probability measure are the Gaussian wave-packets.
-This means that any other non-negative Wigner function will be in the convex hull of Gaussian wave-packets, and therefore will allow a decomposition in terms of them.
+Recall that an ensemble space must be a convex subset of a vector space.
+If we are able to represent ensembles as quasi-probability distribution over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such an $Z$ always exists is still open for us.
+A possible conjecture is that $Z$ could be constructed from the statistical quantities $F_i$ that fully identify each ensemble in $\Ens$.} In this case, any element $v\in V$ can be represented uniquely by a signed probability measure $\mu_{v}$, regardless of whether it is an actual ensemble or not.
+Since statistical quantities $F_i$ are linear functional, we will have $F_i(v) = \int_Z F_i(z) d\mu_{v}$.
+In other words, there is nothing special about quantum mechanics that allows quasi-probability representations.
+It is the fact that the ensemble space embeds into a vector space, a general fact for all theories, that allows such representations.
 
-Note that if $\overline{\Ens} = M_p(X)$, then the space of the extreme point is exactly the space of Dirac measures, and therefore $X$ corresponds to the pure states.
-Alternatively, either one ensemble is represented by an affine measure, or not all probability measures can be understood as mixtures of extreme points.
-If $\overline{\Ens}$ is compact, then this is exactly when $\overline{\Ens}$ is not a simplex and multiple decompositions are allowed.
+It should be clear that, since an ensemble space is a strict subset of the embedding vector space, not all affine measure will correspond to an ensemble space.
+Whether all probability measures correspond to a valid ensemble, or whether signed probability measures are even used depends on how the mapping is done in each particular case.
+For example, there is no simple characterization of which Wigner functions correspond to valid quantum states and which do not.
 
@@
 
-This alternative way of looking at measure theoretic representations, then, is consistent with the previous one.
-While quasiprobability distributions are very useful in practice, we want to stress that there is no single choice for $X$.
-To solve a specific problem, this may actually be useful as one can choose the most convenient $X$ for the problem.
-To develop a general theory for physics, however, this is not useful as each representation is not general enough.
+\textbf{Non-uniquess of quasi-probability.}
+
+This alternative way of looking at measure theoretic representations, then, is consistent with the previous one and fits in our general framework.
+Since quantum ensemble spaces are not simplexes, classical probability does not apply.
+The different bounds one gets between classical and quantum mechanics depend on the different shape of the respective ensemble spaces.
+
+Quasi-probability distributions are very useful in practice, particularly because one can construct them to fit particular problems.
+Once one has identified a set of observables that are the most natural ones to define a problem, one can find a representation that is specifically suited for the case.
+While the quasi-probability cannot be interpreted as probability, it still provides a useful tool to describe and understand the problem.
+
+However, this practical convenience is also a drawback for finding a more general approach to probability in spaces beyond classical mechanics.
+Crucially, the space $Z$ over which the quasi-probability measure are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tells us which quasi-probability measure represent valid ensembles and which do not.
 
@@
 Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
-Note that, given the uniform distribution, $1/6$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
+Note that, given the uniform distribution, $1/6$ is exactly the probably for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
 This gives the basic insight for our definitions.
@@
 The fraction is always well-defined because we can always write $\ens = 0 \ens[a] + 1 \ens$, therefore it must be zero or greater.
-This quantity tells us how much of the ensemble $\ens$ can be constructed from $\ens[a]$.
+This quantity tells us how much of ensemble $\ens$ can be constructed from $\ens[a]$.
 
@@
 Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.
-mixtures),\footnote{Since we work with convex sets within a topology, the hull is both the closure under the topology and convex combinations.
+mixtures),\footnote{Since we work with convex sets within a topology, the hull is both the closed under the topology and convex combinations.
 That is, $\hull(A)$ is the smallest closed and convex set that contains $A$} we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
@@
 For sub-additivity, note that the fraction of two components can at most sum during mixing.
-For the last, increasing and decreasing sequences strictly add or remove possible mixtures; therefore, they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
+For the last, increasing and decreasing sequences strictly add or remove possible mixtures, therefore they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
 The limit of these sequences must agree with the fraction capacity of the limit.
@@
 The fraction capacity, in fact, is not just defined on the extreme points, but it is defined on the whole space of ensembles.
-In the classical discrete case, for example, if $A \in \Omega \ni \Ens$ is an event, then $\frcap_{\ens}(A)$ will return the probability of event $A$; if $P \subset \Ens$ represents all the Poisson distributions with integer coefficient, $\frcap_{\ens}(P)$ will return how much of $\ens$ can be expressed as a mixture of those distributions.
+In the classical discrete case, for example, if $A \subset \Omega \subset \Ens$ is an event, then $\frcap_{\ens}(A)$ will return the probability of event $A$; if $P \subset \Ens$ represents all the Poisson distributions with integer coefficient, $frcap_{\ens}(P)$ will return how much of $\ens$ can be expressed is a mixture of those distributions.
 Similarly, the probability measures in Choquet theory are, in general, defined on the whole ensemble space, and then one finds those particular ones that have support on the extreme points.
@@
 This means that each element of the ensemble space is represented by a unique probability measure over the extreme points.
-In that case, the fraction capacity, extended to the extreme points, will also be additive.
+In that case, the fraction capacity extended on the extreme points will also be additive.
 In the non-classical case, the closure of the ensemble space will not be a simplex, and there will be multiple representations, which will lead to a non-additive measure.
 Suppose, in fact, that $\mu$ and $\nu$ are two distinct measures that represent the same ensemble $\ens$.
-Then there will be a set $A$ such that $\mu(A) \neq \nu(A)$.
+Then there will be a set $A$ such such that $\mu(A) \neq \nu(A)$.
 Suppose, without loss of generality, that $\mu(A) > \nu(A)$ We will also have $\mu(A^{\complement}) = 1- \mu(A) < 1 - \nu(A) = \nu(A^{\complement})$.
@@
 Since entropy characterizes the variability of the preparations of the ensemble, greater variability means the ensemble is spread over more cases.
-Therefore, the count of distinguishable states must be a monotonic function of the entropy.
+Therefore the count of distinguishable states must be a monotonic function of the entropy.
 It has been shown[cite Hall] under very general conditions that the exponential of the entropy is the only function to have the appropriate features.
@@
 The first property readily comes from the range of the exponential.
-For the second, note that $\hull$, exponential, and $\sup$ are all monotone.
+For the second, note that $\hull$, exponential and $\sup$ are all monotone.
 For sub-additivity, the exponential of the entropy itself is sub-additive, and it is additive over orthogonal ensembles.
@@
 
-Again, we note that the non-additive measure we recover is very well-behaved and connects directly with important quantities.
+Again we note that the non-additive measure we recover is very well-behaved, and connects directly with important quantities.
 When Choquet theory applies, state capacity can be extended to the extreme points.
 In the classical case, disjoint sets of extreme points will be orthogonal, meaning that the state capacity there restricted is an additive measure.
-In the discrete case, every point is associated with zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
-In the continuous case, given a subset of phase space $A$, the maximum entropy is reached by the uniform distribution $\rho_A$, which means we recover $\log_2 \capacity(A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
+In the discrete case, every point is associated to zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
+In the continuous case, gives a subset of phase space $A$, the maximum entropy is reached by the uniform distribution $\rho_A$, which means we recover $\log_2 \capacity(A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
 Similarly, in quantum mechanics, given an orthogonal subspace $A$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(A) = 2^{\dim(A)}$.
@@
 In classical mechanics, all states are perfectly distinguishable.
-In quantum mechanics, only orthogonal states are distinguishable, and therefore the state capacity recovers the number of orthogonal states.
+In quantum mechanics, only orthogonal state are distinguishable, and therefore the state capacity recovers the number of orthogonal states.
 

### 2025-07-03 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAddititiveProbability: more work on the standard probability section

@@
 
-\section{Additive measure and Choquet theory}
+\section{Probability measures and Choquet theory}
 
-[This section has to make clear in what extra conditions on ensembles spaces Choquet theory can be applied.]
+\textbf{Classical probability measures.} In the classical case, ensembles are defined as probability measures of the space of all possible states $X$.
+In the discrete case, like for roll of a die, $X$ is a set endowed with the discrete topology\footnote{Physically, the discrete topology corresponds to the ability to experimentally verify or falsify each case.
+That is, we are able to recognize whether a particular die landed or didn't land on a particular face.} and, consequently, the power set as the $\sigma$-algebra.
+In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variable which are verifiable up to finite but arbitrarily small precision.
+That is, we are able to tell whether a state is within a particular range of position and momentum.} and the standard Borel sets as the $\sigma$-algebra.
+In either case, a probability measure is a map $p : \Sigma_X \to [0,1]$ that assign to each event of the $\sigma$-algebra (i.e.
+to an empirical statements) the probability to find it to be true.
+The probability measure is additive, meaning that if two events $A, B \in \Sigma_X$ are mutually exclusive (i.e.
+$A \cap B = \emptyset$) then the probability of the union is the sum of the probability of the respective events:
+\begin{equation}
+	p(A\cup B) = p(A) + p(B).
+\end{equation}
+
+\textbf{Absolute continuity of probability measures.} While each classical ensemble is a probability distribution, not all probability distribution can be thought as a valid ensemble.
+For example, we could have a probability measure to be wholly defined over a single point of phase space, which would correspond to the ability to prepare a system in the same exact place with the same exact momentum.
+Not only this is not feasible experimentally, which violates the whole premise of having objects that map to physically meaningful entities, but it would have to have entropy equal minus infinity.
+To address this, we add a requirement that is both mathematically natural and physically meaningful.
+Let $\mu : \Sigma_X \to [0, + \infty]$ the measure on $X$ that quantifies the count of states in each region.
+In the discrete case this is simply the counting measure (i.e.
+three possible faces of the die corresponds to three possible states), while in the continuous case this corresponds to the Liouville measure (i.e.
+volumes in phase space correspond to the count of states).
+This measure is additive over mutually exclusive events.
+The requirement is that we can assign non-zero probability only to events that have a non-zero count of states.
+That is, if for a particular event $A$ we have $p(A) \neq 0$, then we must have that $\mu(A) \neq 0$ as well.
+Mathematically, $p$ is absolutely continuous with respect to $\mu$, and this is exactly the case in which a probability density $\rho = \frac{dp}{d\mu}$ can be defined.\footnote{In the general case, we can still assign non-zero probability to a point of a manifold, provided that the topology and the measure $\mu$ consistently allow for it.
+Physically, it means that those points are ``special'' both in terms of experimental verifiability and entropy/count of states.
+While the general framework allows for it, it is not crucial for our discussion here.} This is another example in which a physical requirement (probability must be zero if there are no possible states), once properly spelled out (absolute continuity), justifies common physical intuition (we assume there is a probability density).
 
 % Intro paragraph that says quickly how to connect to Choquet theory and its limitations. It should be a quick summary of the whole section
-Since ensemble spaces are a convex subset of a locally convex topological vector space, they can be connected to Choquet theory.
 
-% Quick review of Choquet theory
-Let us first give a summary of the main results in Choquet theory that are interesting for our purposes.
-First of all, let $E$ be a convex subset of a locally convex space $V$.
+\textbf{Choquet theory.} Having reviewed the use of probability measures in classical mechanics, we can ask the more general question: when can we represent ensembles in a generic ensemble space as a probability measure?
+This question is treated by Choquet theory, though in not the full generality that we need it.
+Let us therefore review the main results of Choquet theory, before discussing how it connects to ensembles spaces.
+
+Mathematically, the goal of Choquet theory is to represent a convex set as a probability measure over the extreme points.
+Let $E$ be a convex subset of a locally convex space $V$.
 We say that $\ens \in E$ is represented by a probability measure $\mu : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_E f d\mu$ for every continuous linear functional $f$ on $V$.
@@
 \end{thrm}
+Since we saw that an ensemble space is convex subset of a locally convex topological vector space, the relevance of Choquet theory is straight forward.
 

### 2025-07-04 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: another pass at classical probability

@@
 \textbf{Choquet theory.} Having reviewed the use of probability measures in classical mechanics, we can ask the more general question: when can we represent ensembles in a generic ensemble space as a probability measure?
-This question is treated by Choquet theory, though in not the full generality that we need it.
-Let us therefore review the main results of Choquet theory, before discussing how it connects to ensembles spaces.
+A very similar question is the subject of Choquet theory.
 
-Mathematically, the goal of Choquet theory is to represent a convex set as a probability measure over the extreme points.
-Let $E$ be a convex subset of a locally convex space $V$.
-We say that $\ens \in E$ is represented by a probability measure $\mu : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_E f d\mu$ for every continuous linear functional $f$ on $V$.
-The Choquet theorem states:
+First, we have to define what it means to represent an ensemble with a probability measure.
+As a first step, let $E$ be a convex subset of a locally convex space $V$.
+We say that $\ens \in E$ is represented by a probability measure $p : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_E f dp$ for every continuous linear functional $f$ on $V$.
+Physically, instead of describing our ensemble $\ens$ as a point in our ensemble space identified by the value of all possible statistical variables $f(\ens)$, we represent it as a probability measure $p$ and all the associated expectation values.
+
+In this first step, we are using a probability measure on $E$, so we are defining a probability distribution over ensembles.
+As a second step, we want the probability measure to be defined over the possible cases, the sample space.
+To do that, Choquet theory identifies each possible case with the Dirac measure associated to that case, which is an extreme point of the convex set $E$.
+The result is the following:
 \begin{thrm}[Choquet]
@@
 \end{thrm}
-Since we saw that an ensemble space is convex subset of a locally convex topological vector space, the relevance of Choquet theory is straight forward.
 
-% Physical meaning of representation with a measure.
-First, let us understand the physical significance.
-An ensemble $\ens$ can be seen as a probability measure $\mu$ over the ``pure states'' (i.e.
-the extreme points of $E$) if the value for every statistical quantity $f$ for $\ens$ can be recovered as an expectation over the pure states using the probability measure.
-That is, all statistical properties of $\ens$ can be recovered by the probability measure.
-This is what allows to think go $\ens$ as a probability distribution.
-Note, however, that the Choquet theorem does not guarantee uniqueness.
-The probability measure is unique if and only if $E$ is a Choquet simplex, which is a generalization of the finite dimensional simplex.
-The characterization of the Choquet simplex is fairly technical, so we will simply define the Choque simplex by the property of having a single decomposition in terms of extreme points.
+Note that while an ensemble space $\Ens$ is a metrizable convex subset of a locally convex TVS $V$, it is not necessarily compact.
+As we saw, the Dirac measures should not be considered valid ensembles in the continuum case.
+Therefore Choquet theory cannot be applied to our general case.
+When we are drawing the parallel, we will implicitly assume that $E$ is a compact set that is also the closure of $\Ens$ in an appropriate convex TVS $V$.
+Many interesting cases fall into this restriction: the ensemble spaces of classical probability distribution over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
+When Choquet theory applies, then, we are able to always characterize an ensemble in terms of a probability distributions over its, maybe idealized, most refined ensembles.
 
-% Limitation of applying Choquet theory to our case
-We should now look at the requirements of the theorem, to see whether we can always apply it to an ensemble space.
-First of all, and ensemble space $\Ens$ is a convex subset of a convex sapce $V$.
-Moreover, since every Hausdorff first-countable topological vector space is metrizable, $\Ens$ is metrizable.
-However, $\Ens$ is not necessarily compact.
-In fact, $\Ens$ is not necessarily a closed subset of $V$.
-Consider the space of probability measures over $\mathbb{R}$.
-Since $\mathbb{R}$ is not compact, the space of probability measure over $\mathbb{R}$ cannot be compact.\footnote{TODO: Check with mathematicians} Moreover, the set of extreme points is given by the Dirac measures (i.e.
-probability one over a single point), which would have minus infinite entropy, and therefore cannot be part of the ensemble space.
-Therefore the theorem does not apply in general.
+The previous theorem does not, however, guarantee uniqueness.
+The above theorem, in fact, applies in the quantum case.
+The Bloch ball, in fact, is a compact convex subset and any point can be seen as a convex combination of points on the surface.
+The center, which represent the maximally mixed state, can be thought as an equal mixture of any two opposite points, or a uniform distribution over the whole surface.
+The fact that any ensemble can be seen as a mixture of a set of pure states, then, is not a property of classical spaces.
+The property of classical spaces is that this representation is unique.
 
-% Define "Choquet space" as those spaces where we can apply Choquet theory
-We can apply Choquet theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
-TODO: finalize the proof} and metrizable.
-In this case, we are guaranteed that all ensembles allows a representation through a probability on points that may corresponds idealized ensembles (i.e.
-the Dirac measure) that are not, even conceptually, realizable.
-Many interesting cases fall into this restriction: the ensemble spaces of classical probability distribution over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
-
-Note that ensembles over classical spaces will be represented by a unique probability measure, therefore the corresponding closure of the ensemble space is a Choquet simplex.
-However, quantum mixed states can be decomposed in multiple ways in terms of pure states, and therefore the closure of a quantum ensemble space is not a simplex.
+Mathematically, the probability measure is unique if and only if $E$ is a Choquet simplex, which is a generalization of the finite dimensional simplex.
+The characterization of the Choquet simplex is fairly technical, so we will simply define the Choquet simplex by the property of having a single decomposition in terms of extreme points.
+The classical case, that is when the ensemble space coincide with the space of all probability measures over a set of states, is a Choquet simplex.
 
@@
 
-Chouqet theory tells, in the realm where it can be applied, that classical probability fails in quantum mechanics precisely because the ensemble space of a quantum system allows multiple decomposition.
+The core insight from Chouqet theory, then, is that both in the classical and quantum case we can represent ensembles as a probability distribution over a set of states.
+However, only in the classical case this representation is unique.
+In the quantum case, mixed states can be decomposed into pure states in multiple ways, and it is this multiple decomposition that makes the ensemble space not a simplex, and therefore the space not classical.
+Unique representation of ensembles is a desirable property, and standard probability measures are not going to provide it in general.
 
@@
 Crucially, the space $Z$ over which the quasi-probability measure are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tells us which quasi-probability measure represent valid ensembles and which do not.
+
+As a comparison, in the previous case all probability measures were given over a uniquely defined set of points, the extreme points of the ensemble space, but the probability measure may not be unique.
+In this case, the pseudo-probability associated to each ensemble is unique, but the space over which the pseudo-probability is defined is not.
 

### 2025-07-07 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Add examples: Wigner, Husimi Q, ...
Defined the probability axioms of Kolmogorov.
Gave examples of quasiprobability functions: Wigner, Husimi Q, Glauber P, SIC-POVMs.

--- 2025-07-02 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 \begin{abstract}
-	We present a physically motivated generalization of probability theory that is suitable for classical mechanics, quantum mechanics and any future physical theory that allows a statistical description.
-The goal is to put the use of classical and quantum probability in a broader context, and show how the current mathematical structures are likely not suitable to solve the open problems in the foundations of physics.
-For the more mathematically inclined, we will point to areas where new math or generalization of established math are needed.
-For the more philosophically incline, we will point to areas where further conceptual work is needed.
+	We present a physically motivated generalization of probability theory that is suitable for classical mechanics, quantum mechanics, and any future physical theory that allows a statistical description.
+The goal is to put the use of classical and quantum probability in a broader context and show how the current mathematical structures are likely not suitable to solve the open problems in the foundations of physics.
+For the more mathematically inclined, we will point to areas where new math or generalization of established math is needed.
+For the more philosophically inclined, we will point to areas where further conceptual work is needed.
 
@@
 The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
-From the entropy, we can define the state capacity, which generalizes the notion of count of states.
+From the entropy, we can define the state capacity, which generalizes the notion of the count of states.
 For each ensemble, we can define the fraction capacity, which generalizes the notion of probability.
 Both are non-negative sub-additive monotonic set functions, with the second being unit bounded.
-These measures are always unique and, since they are defined on the ensemble space itself, always have provide a physically meaningful representation.
-Moreover, in the compact case, these measure can be restricted on the extreme points while retaining uniqueness.
+These measures are always unique and, since they are defined on the ensemble space itself, always provide a physically meaningful representation.
+Moreover, in the compact case, these measures can be restricted to the extreme points while retaining uniqueness.
 (TODO: prove)
@@
 
+A probability measure $\mu: \Omega_X \to \mathbb{R}$ over a sample space $X$ with $\sigma$-algebra $\Omega_X$ satisfies the three axioms by Kolmogorov
+\begin{enumerate}[label={\Roman*}]
+	\item\label{ax:pos} \textbf{Non-negativity:\,} $\forall x \in \Omega_X: \mu(x) \ge 0$
+	\item\label{ax:norm} \textbf{Normalization:\,} $\mu(X) = 1$
+	\item\label{ax:add} \textbf{$\sigma$-additivity:\,} $\mu ( \bigcup_i x_i ) = \sum_i \mu ( x_i )$ for any countable sequence of disjoint sets $x_i \in \Omega_X$.
+\end{enumerate}
+The last axiom is based on the (sometimes hidden) assumption of
+\begin{enumerate}[label={\Roman*}]\setcounter{enumi}{3}
+	\item\label{ax:exc} \textbf{Mutual exclusivity:\,} disjoint sets describe mutually exclusive events.
+\end{enumerate}
+
 % Physical meaning of representation with a measure.
@@
 Consider the space of probability measures over $\mathbb{R}$.
-Since $\mathbb{R}$ is not compact, the space of probability measure over $\mathbb{R}$ cannot be compact.\footnote{TODO: Check with mathematicians} Moreover, the set of extreme points is given by the Dirac measures (i.e.
+Since $\mathbb{R}$ is not compact, the space of probability measures over $\mathbb{R}$ cannot be compact.\footnote{TODO: Check with mathematicians} Moreover, the set of extreme points is given by the Dirac measures (i.e.
 probability one over a single point), which would have minus infinite entropy, and therefore cannot be part of the ensemble space.
-Therefore the theorem does not apply in general.
+Therefore, the theorem does not apply in general.
 
 % Define "Choquet space" as those spaces where we can apply Choquet theory
-We can apply Choquet theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
+We can apply Choquet's theorem in the case where the closure of $\Ens$ in $V$ is a compact set $E$, which will also be convex\footnote{The closure of a convex set is convex.
 TODO: finalize the proof} and metrizable.
-In this case, we are guaranteed that all ensembles allows a representation through a probability on points that may corresponds idealized ensembles (i.e.
+In this case, we are guaranteed that all ensembles allow a representation through a probability on points that may correspond to idealized ensembles (i.e.
 the Dirac measure) that are not, even conceptually, realizable.
-Many interesting cases fall into this restriction: the ensemble spaces of classical probability distribution over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
+Many interesting cases fall into this restriction: the ensemble spaces of classical probability distributions over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
 
 Note that ensembles over classical spaces will be represented by a unique probability measure, therefore the corresponding closure of the ensemble space is a Choquet simplex.
-However, quantum mixed states can be decomposed in multiple ways in terms of pure states, and therefore the closure of a quantum ensemble space is not a simplex.
+However, quantum mixed states can be decomposed in multiple ways in terms of pure states, and therefore, the closure of a quantum ensemble space is not a simplex.
 
@@
 
-%Things to say: some people think about it as weakining axioms (cite papers on negative probability).
+The Wigner function is one of the most well-known examples of a quasi-probability function. The Wigner transform
+\begin{equation}
+	W_{\hat{O}} (q, p) = 2 \int_{-\infty}^{\infty} \langle x + y | \hat{\rho} | x - y \rangle e^{-2 i p y / \hbar} \; \mathrm dy
+\end{equation}
+maps operators (i.e., states and observables) on the Hilbert space onto functions on phase space.
+It allows calculation of expectation values as expected by turning the trace into a phase space integral
+\begin{equation}
+	\tr ( \hat{\rho} \hat{O} ) = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} W_{\hat{\rho}} (q, p) W_{\hat{O}} (q, p) \; \mathrm dq \; \mathrm dp \; .
+\end{equation}
+This representation is most convenient when working with operators in symmetric ordering.
+
+However, these functions don't define probability densities for two reasons.
+First of all, they can be negative on small regions\footnote{compared to $\hbar$} of phase space.
+And second of all, the events\footnote{technically measurable subsets like $[q - \epsilon, q + \epsilon] \times [p - \delta, p + \delta]$, etc.} $(q, p)$ are not mutually exclusive, which can be seen by the fact that Wigner functions of orthogonal pure states can overlap.
+For example, the eigenstates $| m \rangle$ of the harmonic oscillator with Wigner functions
+\begin{equation}
+	W_m (q, p) = \frac{(-1)^m}{\pi} e^{-(q^2+p^2)} L_m (2(q^2 + p^2))
+\end{equation}
+have support everywhere, since the Laguerre polynomials $L_m$ have finitely many zeros.
+
+The converse, however, is true, i.e., Wigner functions of disjoint support\footnote{TODO It is not obvious if any physical examples exist.
+One could attempt to construct density operators from two real Schwartz functions on $\mathbb{R}^2$ with disjoint support, which would yield self-adjoint trace-class operators, but they aren't necessarily positive.} are orthogonal, since the inner product of two pure states $\psi, \phi$ can be calculated as
+\begin{equation}
+	| \langle \psi | \phi \rangle |^2 = 2 \pi \hbar \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} W_\psi (q, p) W_\phi (q, p) \; \mathrm dq \; \mathrm dp \; .
+\end{equation}
+
+The Husimi Q function is another example of a quasi-probability density. It is convenient when working with operators in anti-normal order and defined as
+\begin{equation}\label{eq:husimi}
+	Q ( \alpha ) = \frac{1}{\pi} \langle \alpha | \hat{\rho} | \alpha \rangle
+\end{equation}
+where the $| \alpha \rangle$ are coherent states.
+
+Interestingly, these functions are always non-negative, therefore satisfying axiom \ref{ax:pos}.
+But probability densities are still not related to mutually exclusive events, as demanded by assumption \ref{ax:exc}.
+
+%Things to say: some people think about it as weakening axioms (cite papers on negative probability).
 It's more than that.
 Points describe events that are not necessarily mutually exclusive.
-TODO: example of two Wigner function/Husimi Q would be better that have no overlap but they are not orthogonal and/or two functions that have overlap but are orthogonal.
+TODO: example of two Wigner functions/Husimi Q would be better, that have no overlap but they are not orthogonal and/or two functions that have overlap but are orthogonal.
+
+For example the eigenstates $| n \rangle$ of the harmonic oscillator are orthogonal, but overlap with any coherent state $| \alpha \rangle$ for $\alpha \neq 0$.
+Their Husimi Q function $Q_n ( \alpha ) = \frac{1}{\pi} \langle \alpha | \hat{\rho_n} | \alpha \rangle = \frac{1}{\pi} \langle \alpha | n \rangle \langle | n | \alpha \rangle = \frac{1}{\pi} | \langle n | \alpha \rangle |^2 = \frac{1}{\pi} e^{| \alpha |^2} \frac{\alpha^{2n}}{n!} $ have identical support on all of $\mathbb{C}$\footnote{TODO We might be able to prove that Husimi Q functions always have support on all of $\mathbb{C}$, by expressing them via the \href{https://en.wikipedia.org/wiki/Segal\%E2\%80\%93Bargmann_space#The_Segal\%E2\%80\%93Bargmann_transform}{Segal-Bargmann transform}, which is a holomorphic function and as such has support everywhere.
+I'm not sure if there is a way to extend this to Wigner functions as well.}.
+
+The opposite case, however, can never happen.
+Pure states with Husimi Q functions of disjoint support must be orthogonal.
+To see that remember that for pure states $| \psi \rangle$ the corresponding Husimi Q function is $Q_\psi ( \alpha ) = \frac{1}{\pi} | \langle \psi | \alpha \rangle |^2$ and the coherent states $| \alpha \rangle$ form an overcomplete basis, i.e.
+give a partition of unity $1 = \frac{1}{\pi} \int | \alpha \rangle \langle \alpha | \; \mathrm d^2 \alpha$.
+That means we can write the inner product between two pure states $\psi, \phi$ as
+\begin{equation}
+	\langle \psi | \phi \rangle = \frac{1}{\pi} \int \langle \psi | \alpha \rangle \langle \alpha | \phi \rangle
+\end{equation}
+If $Q_\psi$ and $Q_\phi$ have disjoint support for each $\alpha$, either $\langle \psi | \alpha \rangle$ or $\langle \alpha | \phi \rangle$ vanish.
+
+Many other examples exist, like the Glauber-Sudarshan P function, which is related to normal ordering.
+In the finite-dimensional case, one approach is to expand the density operator in terms of symmetric, informationally-complete, positive operator-valued measures (SIC-POVMs)\footnote{The exact definition is not relevant for this article and can be found in the literature.}.
+These can be constructed, for example, from a set of $d^2$ rank-1 projectors $P_i$ and there corresponding probabilities $p_i$ as
+\begin{equation}
+	\hat{\rho} = \sum_{n=1}^{d^2} \left[ (d^2 + 1) p_i - \frac{1}{d} \right] P_i \; ,
+\end{equation}
+where $p_i = \tr P_i \hat{\rho}$. There is no unique choice of projectors for a given Hilbert space and it is in fact an open question if SIC-POVMs exist in all dimensions $d$.
 
@@
 For example, the Husimi Q representation provides us with a non-negative quasi-probability function.
-That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.
-Similarly, since we can embed the Bloch ball inside a tetrahedron, we can represent all states of a two-state quantum system by a convex combination of the vertices, leading to a non-negative quasi-probability distribution over four elements.
+That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a probability distribution.
 
-Conversely, we can use signed measure to represent classical distributions.
-Suppose we have a two state classical discrete space, which means the space of all possible ensembles is spanned by $[p, (1-p)]$.
-Alternatively, we can write it as $p[1,0] + (1-p)[0,1]$ which highlights that we are writing each possible probability distribution in terms of the two possible cases.
-But we could also express each probability distribution in terms of other two distributions, as in $\left(2p - \frac{1}{2}\right)\left[\frac{3}{4}, \frac{1}{4}\right] + \left(\frac{3}{2} - 2p\right)\left[\frac{1}{4}, \frac{3}{4}\right]$.
-The coefficients now span between $\left[-\frac{1}{2}, \frac{3}{2}\right]$, can be negative, but still sum to one.
-That is, in terms of this new affine basis, we need to use a signed probability.
-
-Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measure, the difference between classical probability and quantum quasi-probability does not lie in the use of negative probability.
-Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events.
-That is, they can be prepared and observed independently from the others; we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasi-probability in quantum mechanics.
+The crucial difference is that in classical probability all points of the sample space represent mutually exclusive events.
+That is, they can be prepared and observed independently from the others.
+Therefore, any two events that are disjoint are also mutually exclusive.
+This is not the case for quasi-probability in quantum mechanics.
 Husimi Q represents a state as decomposed over coherent states, which cannot represent mutually exclusive as they are not orthogonal.
-The inner product between these states will provide a constrain.
-In the non-negative probability representation of the Bloch sphere, none of the four corners represent valid states, as they lie outside the Bloch ball.
-So we cannot assign probability to a vertex without assigning some probability to another vertex.
-
-While negative probability is a red herring, we want to understand two things.
-When can we use classical probability?
-Are quasi-probability representation specific to quantum mechanics, or are they more general?
-Fortunately, the ensemble space framework allows us to answer these questions directly.
+Similarly, we can imagine to embed the Bloch ball into a tetrahedron, which means every point in the Bloch ball could be represented as a convex combination of extreme points, meaning that they would be represented by a non-negative quasi-probability distribution over four elements.
+However, none of the four elements represent a valid state.
 
@@
 
-\textbf{Ensemble space embedding and signed measures.} We have already seen that we can represent ensembles using classical probability if and only the ensemble space is a simplex.
-Since any probability measure is possible, we can assign probability to two disjoint areas independently.
-This answers the first question.
+For example, affine measures do not necessarily satisfy non-negativity. The Wigner function is the probability density of such a measure and can be negative over small regions of phase space.
 
-For the second question, let $Z$ be the space over which we are defining our quasi-probability distributions.
-Note that, in general, $Z$ is not the space of states.
-Wigner functions, for example, are defined over phase space, which does not represent quantum states.
-Let $M^{\pm}(Z)$ the space of all bounded signed measures over $Z$.
-Since the linear combination of two bounded signed measure is a bounded signed measure, $M^{\pm}(Z)$ is a vector space.
-Now let $M_1^{\pm}(Z)$ be the space of signed probability measures.
-While it is not a subspace of $M(Z)$, it does not include the zero measure, it is an affine subspace therefore, up to the choice of an arbitrary origin, it is a vector space.
+Given a space $X$, the space of all affine measures is a vector space, not just a convex subset as in the case of the space of probability measures.
+This means that, given an ensemble space, we may find a space $X$ such that the space of affine measures over $X$ is isomorphic to the embedding vector space $V$ as topological vector spaces.
+In this case, any element $v\in V$ can be represented uniquely by an affine measure $\mu_{v}$, regardless of whether they are actual ensembles or not.\footnote{The question of whether such an $X$ always exists is left open.
+We suspect a link exists between the statistical quantities $F_i$ that fully identify each ensemble and the space $X$.} Because the statistical quantities $F_i$ are affine functions, we will have $F_i(v) = \int_X F_i d\mu_{v}$.
 
-Recall that an ensemble space must be a convex subset of a vector space.
-If we are able to represent ensembles as quasi-probability distribution over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such an $Z$ always exists is still open for us.
-A possible conjecture is that $Z$ could be constructed from the statistical quantities $F_i$ that fully identify each ensemble in $\Ens$.} In this case, any element $v\in V$ can be represented uniquely by a signed probability measure $\mu_{v}$, regardless of whether it is an actual ensemble or not.
-Since statistical quantities $F_i$ are linear functional, we will have $F_i(v) = \int_Z F_i(z) d\mu_{v}$.
-In other words, there is nothing special about quantum mechanics that allows quasi-probability representations.
-It is the fact that the ensemble space embeds into a vector space, a general fact for all theories, that allows such representations.
+In general, since the ensemble space is bounded along every direction, not all affine measures will correspond to an ensemble space.
+Whether all probability measures correspond to a valid ensemble, or whether affine measures are even used, corresponds to relationships between $\Ens$ and the space of probability measures $M_p(X)$.
+If $\Ens \subset M_p(X)$, then all states are represented by a probability measure.
+However, this is not a ``probability'' in the standard sense of the word.
+For example, the Husimi Q representation is always non-negative at every point, but two points do not represent mutually exclusive events as they are not orthogonal (i.e.
+their inner product is non-zero and they do not maximize the entropy increase during mixture).
+Mathematically, the issue is that the points $X$ of the Husimi Q representation are not the extreme points of the closure ensemble space $\overline{\Ens}$.
+In general, $\Ens$ and $M_p(X)$ will overlap, but neither will contain the other.
+For example, in the Wigner representation, the only pure states (i.e.
+the extreme points of $\Ens$) represented by a probability measure are the Gaussian wave-packets.
+This means that any other non-negative Wigner function will be in the convex hull of Gaussian wave-packets, and therefore will allow a decomposition in terms of them.
 
-It should be clear that, since an ensemble space is a strict subset of the embedding vector space, not all affine measure will correspond to an ensemble space.
-Whether all probability measures correspond to a valid ensemble, or whether signed probability measures are even used depends on how the mapping is done in each particular case.
-For example, there is no simple characterization of which Wigner functions correspond to valid quantum states and which do not.
+Note that if $\overline{\Ens} = M_p(X)$, then the space of the extreme point is exactly the space of Dirac measures, and therefore $X$ corresponds to the pure states.
+Alternatively, either one ensemble is represented by an affine measure, or not all probability measures can be understood as mixtures of extreme points.
+If $\overline{\Ens}$ is compact, then this is exactly when $\overline{\Ens}$ is not a simplex and multiple decompositions are allowed.
 
@@
 
-\textbf{Non-uniquess of quasi-probability.}
-
-This alternative way of looking at measure theoretic representations, then, is consistent with the previous one and fits in our general framework.
-Since quantum ensemble spaces are not simplexes, classical probability does not apply.
-The different bounds one gets between classical and quantum mechanics depend on the different shape of the respective ensemble spaces.
-
-Quasi-probability distributions are very useful in practice, particularly because one can construct them to fit particular problems.
-Once one has identified a set of observables that are the most natural ones to define a problem, one can find a representation that is specifically suited for the case.
-While the quasi-probability cannot be interpreted as probability, it still provides a useful tool to describe and understand the problem.
-
-However, this practical convenience is also a drawback for finding a more general approach to probability in spaces beyond classical mechanics.
-Crucially, the space $Z$ over which the quasi-probability measure are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tells us which quasi-probability measure represent valid ensembles and which do not.
+This alternative way of looking at measure theoretic representations, then, is consistent with the previous one.
+While quasiprobability distributions are very useful in practice, we want to stress that there is no single choice for $X$.
+To solve a specific problem, this may actually be useful as one can choose the most convenient $X$ for the problem.
+To develop a general theory for physics, however, this is not useful as each representation is not general enough.
 

### 2025-07-07 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex

--- 2025-07-04 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 \section{Probability measures and Choquet theory}
+
+%[This section has to make clear in what extra conditions on ensemble spaces Choquet theory can be applied.]
 
@@
 
+A probability measure $\mu: \Omega_X \to \mathbb{R}$ over a sample space $X$ with $\sigma$-algebra $\Omega_X$ satisfies the three axioms by Kolmogorov
+\begin{enumerate}[label={\Roman*}]
+	\item\label{ax:pos} \textbf{Non-negativity:\,} $\forall x \in \Omega_X: \mu(x) \ge 0$
+	\item\label{ax:norm} \textbf{Normalization:\,} $\mu(X) = 1$
+	\item\label{ax:add} \textbf{$\sigma$-additivity:\,} $\mu ( \bigcup_i x_i ) = \sum_i \mu ( x_i )$ for any countable sequence of disjoint sets $x_i \in \Omega_X$.
+\end{enumerate}
+The last axiom is based on the (sometimes hidden) assumption of
+\begin{enumerate}[label={\Roman*}]\setcounter{enumi}{3}
+	\item\label{ax:exc} \textbf{Mutual exclusivity:\,} disjoint sets describe mutually exclusive events.
+\end{enumerate}
+
 Note that while an ensemble space $\Ens$ is a metrizable convex subset of a locally convex TVS $V$, it is not necessarily compact.

### 2025-07-08 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed ensemble space section

--- 2025-07-04 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-Conceptually, a statistical ensemble is a collection of outputs, taken at once, of a reproducible preparation procedure for a physical system.
-In the classical case, this corresponds to a probability distribution over a discrete set of outcomes (i.e.
-a fair coin) or over phase space (i.e.
+\textbf{Basic requirements for an ensemble space}.
+It is a cardinal rule in Physical Mathematics that all axioms and mathematical definitions must be grounded in physical requirements.
+This way we can develop mathematical frameworks that are physically well founded.
+As part of our larger project, we are developing a general theory of ensemble spaces, the full treatment of which is beyond the scope of this work.
+We limit ourselves to give the initial premises and the final mathematical characterization, without justifications and proofs.\footnote{More details are available in our ongoing work (cite)}
+
+We start by assuming that physical laws deal with experimentally reproducible relationships, which means that they must describe relationships between statistical ensembles.
+Conceptually, a statistical ensemble is the collection of outputs, taken at once, of a reproducible preparation procedure for a physical system.
+In the classical case, this corresponds to a probability distribution over a discrete set of outcomes (e.g.
+a fair coin) or over phase space (e.g.
 the symplectic manifold representing position and momentum for a classical particle).
-In the quantum case, this corresponds to a mixed state (i.e.
+In the quantum case, this corresponds to a mixed state (e.g.
 a positive semi-definite self-adjoint operator of trace one).
 An ensemble space is the collection of all possible ensembles defined in a physical theory, together with a basic mathematical structure that all physical theories must possess.
-The development of a full theory of ensembles goes beyond the scope of this work, and we refer to the appendix and our ongoing work (cite).
 
-For the purpose of this paper, we will assume an ensemble space $\Ens$ is a convex subset of a Hausdorff locally convex topological vector space $V$.
-The topology is the physically natural topology, where each open set represents a verifiable statement.
-The convex structure represents statistical mixing.
-For example, $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture of two ensembles $\ens[a]$ and $\ens[b]$ with probability $p$ and $\bar{p} = 1 - p$.
-Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i}^{\infty} p_i \ens_i$ converges in the space or not.
-We also assume that there is a countable family of affine continuous functions $F_i : \Ens \to \mathbb{R}$ that represent statistical quantities and fully identify each ensemble.
-That is, $F_i(p \ens[a] + \bar{p} \ens[b]) = p F_i(\ens[a]) + \bar{p} F_i(\ens[b])$ and $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.\footnote{It is an open question whether all ensemble spaces must possess this property.} For example, we can imagine $F_i$ representing the expectation of all polynomials of position and momentum.\footnote{This requires all probability distributions to converge to zero at infinity faster than any polynomial.} Mathematically, these statistical quantities can be used to define semi-norms on the space, giving us a locally convex topology.
+An ensemble space $\Ens$ will need to satisfy a three basic requirements.
+First, it needs to guarantee experimental verifiability.
+Mathematically, the ensemble space must be a $\mathsf{T}_0$ second countable topological space, where each open set represents an experimentally verifiable statement.\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Second, it needs to allow statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.
+Mathematically, this constrains the ensemble space to be a convex set.\footnote{Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i}^{\infty} p_i \ens_i$ converges in the space or not.} Third, an entropy function must be defined that characterizes the variability of the elements within an ensemble.
 
-Additional, each ensemble space $\Ens$ is equipped with a continuous entropy function $S : \Ens \to \mathbb{R}$. The entropy of a statistical mixture satisfies the following bounds
+Additionally, we will require that ensembles can be identified through real valued statistical quantities.
+\footnote{It is an open question whether all ensemble spaces must possess this property.
+In a previous work we showed that the dense linear ordering require highly idealized assumptions.} A statistical quantity represents the expectation of a real valued quantity, and is therefore an affine continuous functions $F : \Ens \to \mathbb{R}$.
+That is, $F(p \ens[a] + \bar{p} \ens[b]) = p F(\ens[a]) + \bar{p} F(\ens[b])$.
+The last requirement is that there exists a family $F_i$ of statistical quantities that fully identify each ensemble.
+That is, $F_i(p \ens[a] + \bar{p} \ens[b]) = p F_i(\ens[a]) + \bar{p} F_i(\ens[b])$ and $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.
+
+These requirements interact with each other leading to a rich mathematical structure that connect to many of the structures we use in mathematical physics.
+In this paper, we will examine only connections to measure theoretic representations.
+
+\textbf{Ensemble spaces as convex sets.} In this work, an ensemble space $\Ens$ is a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{The entropy constraints forces the topology to be $\mathsf{T}_1$ the convex space to embed into a vector space.
+A second countable $\mathsf{T}_1$ TVS is metrizable.
+The family of statistical quantities induce a family of semi-norms on the space, which must be countable due to second countability, which makes the TVS locally convex.} The entropy is function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
 \begin{equation}
@@
 \end{equation}
-where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{In the larger work, we can show that the Shannon entropy is the only upper bound possible.} The bounds come from assuming that the entropy quantifies the variability of each ensemble.
-Intuitively, the variability cannot decrease during mixing, and will have an upper bound when mixing two ensembles that are non-overlapping, or orthogonal.
-In fact, we define two ensembles $\ens[a] \ortho \ens[b]$ to be orthogonal when they mixture gives a maximal entropy increase.
-This does recover the notion of orthogonality in the corresponding classical and quantum vector spaces of statistical ensembles.
-Lastly, mixtures preserve orthogonality in the sense that mixing non-overlapping ensembles still gives non-overlapping ensembles.
-More precisely, $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$.
+where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{In the larger work, we can show that the Shannon entropy is the only upper bound possible.}  Intuitively, the variability (i.e.
+the entropy) cannot decrease during mixing, and will have an upper bound when mixing two ensembles whose instances are always distinguishable.
 
-To summarize, the above mathematical structure represents the requirements that ensembles must be experimentally well-defined, allow statistical mixture, can be identified with statistical quantities and provide a well-defined entropy.\footnote{While the full physical justification of the above mathematical structure is beyond the scope of this work, it is important to us that such justification exists and this physically straight forward.
-This is the essence of what we call Physical Mathematics.
-We will summarize some of these justifications in the appendix, leaving the full and more updated discussion in our work in progress.} We now turn to the question of how we can represent both the entropy and each ensemble in terms of some measure theoretic structure.
+\textbf{Orthogonality and separatedness.} We will say that two ensembles are orthogonal, noted $\ens[a] \ortho \ens[b]$, when their mixture gives a maximal entropy increase.
+This will recover the notion of orthogonality for both classical and quantum spaces.
+Another notion of distinction between two ensembles is in terms of the mixtures: two ensembles are separate, noted $\ens[a] \separate \ens[b]$, they cannot be expressed as a mixture of a common ensemble.
+That is, there is no $\ens[c]$ such that $\ens[a] = \alpha \ens[c] + \bar{\alpha} \ens[d]$ and $\ens[b] = \beta \ens[c] + \bar{\beta} \ens$ from some $\alpha, \beta \in (0,1]$ and $\ens[d],\ens \in \Ens$.
+In classical spaces, these two notions coincide: two ensembles are orthgonal and separate if and only if the support of the respective probability distributions overlap on a non-measure-zero set.
+In quantum mechanics, these two notions are different: any two pure states are separate but are not orthogonal.
+
+To summarize, a repeatedly verifiable physical theory requires ensembles, which must be defined experimentally, allow statistical mixing and provide an entropy.
+Additionally, we require the existence of enough statistical quantities to identify each ensemble.
+Under these assumptions, an ensemble space $\Ens$ is a convex set of a metrizable second countable locally convex topological vector space $V$.
+We now turn to the question of how we can represent both the entropy and each ensemble in terms of some measure theoretic structure.
 

### 2025-07-09 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: expanding with examples

@@
 
-\textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions.
-While the Wigner function is probably the most widely known, there are in fact several different approach.
-The Wigner function represent quantum states, both pure and mixed, as a density function over position and momentum.
-This function integrates to one over the full space, like a probability density, but can have small negative regions, unlike a probability density.
+\textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions, of which the Wigner function is probably the most widely known.
+The Wigner transform $W : \Ens \to L^1(\mathbb{R}^2, \mathbb{R}) $, defined as [CITATION NEEDED]
+\begin{equation}
+	W(\rho) = W_{\hat{\rho}} (q, p) = 2 \int_{-\infty}^{\infty} \langle x + y | \rho | x - y \rangle e^{-2 i p y / \hbar} \; \mathrm dy,
+\end{equation}
+takes a mixed state $\rho$ and returns a function of phase space that integrates to one, like a probability density, but can have small negative regions.
 Recall that an affine combination is a linear combination where all coefficients sum to one but are not necessarily non-negative.
 Therefore, in general, quasi-probability distributions can be understood as affine combinations.
+Remarkably, expectation values of polynomial of position and momentum in symmetric ordering, coincide with expectations over the Wigner function.
+For example
+\begin{equation}
+	\tr \left( \frac{1}{2}[XP+PX] \rho \right) = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} xp \; W_{\rho} (q, p) dq dp \; .
+\end{equation}
+Moreover, the transform is an affine map: $W(p \rho_1 + \bar{p} \rho_2 ) = p W(\rho_1) + \bar{p} W(\rho_2)$.
 
-For the Wigner function, the expectation of polynomials of position and momentum over phase space recovers expectations of polynomials of position and momentum in symmetric order (?).
-(Definitions and features?) In some problems, especially in quantum optics, normal ordering and anti-normal ordering are more useful.
+In some problems, especially in quantum optics, normal ordering and anti-normal ordering are more useful.
 The Glauber P representation and the Husimi Q representation work like the Wigner function except that the expectations of polynomials recover the respective order.
@@
 Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
-For example, the SIC-POVM represents density matrices as affine combinations of pure states that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
+For example, the SIC-POVM represents quantum states as affine combinations of pure states $P_i$ that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
+Each ensemble $\rho$ can then be represented as
+\begin{equation}
+	\rho = \sum_{n=1}^{d^2} \lambda_i P_i = \sum_{n=1}^{d^2} \left[ (d^2 + 1) p_i - \frac{1}{d} \right] P_i \; ,
+\end{equation}
+where $p_i = \tr \left(P_i \rho\right)$. [CITATION NEEDED]
 
@@
 All representations, though, have the following features in common.
-First, we have a map $W : \Ens \to Q(Z,\mathbb{R})$ that given each mixed state $\rho \in \Ens$ returns the corresponding quasi-probability function $w(z) \in Q(Z,\mathbb{R})$, where $Q(Z,\mathbb{R})$ is the function space used by the specific representation.
-The representation preserves convex combinations: $W(\sum_i p_i \rho_i) = \sum_i p_i W(\rho_i)$.
+First, we have a map $W : \Ens \to Q(Z,\mathbb{R})$ that given each mixed state $\rho \in \Ens$ returns the corresponding quasi-probability function $w(z) \in Q(Z,\mathbb{R})$, where $Q(Z,\mathbb{R}) \subset L^1(Z,\mathbb{R})$ is the function space used by the specific representation.
+The space $Z$, in fact, can change from representation to representation: it is $\mathbb{R}^{2n}$ for a Wigner function, $\mathbb{C}^n$ for the Husimi Q and $[1, d^2] \subset \mathbb{N}$ for the SIC-POVM.
+All representations preserves convex combinations: $W(\sum_i p_i \rho_i) = \sum_i p_i W(\rho_i)$.
 That is, a statistical mixture of quasi-probability distributions represents the statistical mixture of the corresponding states.
@@
 
-\textbf{Not about negative probability.} It is very common to conceive the main difference between classical probability and these quasi-probabilities as the inclusion of negative probability.
-Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
+\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasi-probability.
+[CITATION NEEDED] Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
 This is a mischaracterization.
 For example, the Husimi Q representation provides us with a non-negative quasi-probability function.
-That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.
-Similarly, since we can embed the Bloch ball inside a tetrahedron, we can represent all states of a two-state quantum system by a convex combination of the vertices, leading to a non-negative quasi-probability distribution over four elements.
+That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.\footnote{Since the Husimi Q is defined as $Q_{\rho} ( \alpha ) = \frac{1}{\pi} \langle \alpha | \rho | \alpha \rangle$, it is proportional to the probability of measuring a coherent (i.e.
+Gaussian state) $| \alpha \rangle$, which is non-negative.
+} Similarly, a SIC-POVM characterizes the Bloch ball by four pure states that form a tetrahedron inscribed in the ball.
+Since we can also inscribe the Bloch ball inside a tetrahedron, we can represent all states of a two-state quantum system by a convex combination of the vertices, leading to a non-negative quasi-probability distribution over four elements.
 
-Conversely, we can use signed measure to represent classical distributions.
+Conversely, we can use signed measures to represent classical distributions.
 Suppose we have a two state classical discrete space, which means the space of all possible ensembles is spanned by $[p, (1-p)]$.
@@
 Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measure, the difference between classical probability and quantum quasi-probability does not lie in the use of negative probability.
-Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events.
-That is, they can be prepared and observed independently from the others; we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasi-probability in quantum mechanics.
-Husimi Q represents a state as decomposed over coherent states, which cannot represent mutually exclusive as they are not orthogonal.
-The inner product between these states will provide a constrain.
-In the non-negative probability representation of the Bloch sphere, none of the four corners represent valid states, as they lie outside the Bloch ball.
-So we cannot assign probability to a vertex without assigning some probability to another vertex.
+Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events, and all mutually exclusive events are represented by disjoint regions.
+Mutually exclusive events and be prepared and observed independently from the others (i.e.
+they are orthogonal in the sense that they maximize entropy increase during mixture); we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasi-probability in quantum mechanics.
+
+For example, the eigenstates $| m \rangle$ of the harmonic oscillator are orthogonal, mutually exclusive events. Their corresponding Wigner functions are
+\begin{equation}
+	W_m (q, p) = \frac{(-1)^m}{\pi} e^{-(q^2+p^2)} L_m (2(q^2 + p^2)).
+\end{equation}
+Their support is the whole $\mathbb{R}^2$ since the Laguerre polynomials $L_m$ have finitely many zeros, and therefore are not disjoint.
+Conversely, in a SIC-POVM we take $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot be all orthogonal to each other.
+For a Bloch sphere, the four corners of the tetrahedron are not all opposite to each other.
+Therefore, disjoint sets of $P_i$ do not correspond to mutually exclusive events.
 

### 2025-07-10 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed fraction capacity

@@
 
-TODO: intro
+\textbf{Probability of outcomes vs fraction of components.} From the previous discussion, we can see that the problem with using probability in quantum mechanics is that not all pure states represent mutually exclusive events, and it is only over mutually exclusive events that classical probability can be defined.
+Mutually exclusive events correspond to ensemble are orthogonal, as each instance of one ensemble cannot be produced by the other ensemble.
+That is why, in quantum mechanics, probability can be defined only once a context, an orthogonal basis, is chosen.
+Therefore, at a conceptual level, probabilities can only give us a patchwork representation of ensembles.
+Some author embrace this view, by characterizing states as the collection of all probability distributions of all possible results, as it is generally the case in GPTs.
+[CITATION NEEDED] However, there is a different route.
 
-Given that mixing coefficients have a more general applicability than probability of outcomes, let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
+Note that the linearity in our axioms is in terms of mixing coefficients, not probability of outcomes.
+That is, regardless of whether we can characterize experiment outcomes with probability distributions, preparation procedures will always allow us to create statistical mixture.
+We can create statistical mixtures regardless of whether the states we are mixing represent mutually exclusive outcomes or not.
+That is, instead of focusing on the likelihood of an outcome, we will focus on how much of a particular ensemble can be understood as the mixture of some other ensembles.
+Instead of focusing on what an ensemble may cause, we focus on how an ensemble can be produced.
+It turns out, this second approach generalizes in a natural way.
+
+\textbf{Fraction capacity.} Let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
 How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
@@
 \end{equation}
-This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$.\footnote{The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
-fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.
-TODO \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} }
+This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$.
 
-One can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
+\textbf{Non-additivity of the fraction capacity.} One can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
 \begin{enumerate}
@@
 The limit of these sequences must agree with the fraction capacity of the limit.
-If not for the lack of additivity over disjoint sets, the fraction capacity would be a probability measure.
 
-Our setup is very similar to what happens in Choquet theory.
-The fraction capacity, in fact, is not just defined on the extreme points, but it is defined on the whole space of ensembles.
-In the classical discrete case, for example, if $A \subset \Omega \subset \Ens$ is an event, then $\frcap_{\ens}(A)$ will return the probability of event $A$; if $P \subset \Ens$ represents all the Poisson distributions with integer coefficient, $frcap_{\ens}(P)$ will return how much of $\ens$ can be expressed is a mixture of those distributions.
-Similarly, the probability measures in Choquet theory are, in general, defined on the whole ensemble space, and then one finds those particular ones that have support on the extreme points.
+Note that, if not for the lack of additivity over disjoint sets, the fraction capacity would be a probability measure.
+It is therefore a non-additive probability measure.
+The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
+fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\footnote{ \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} } We will see how our approach fits  in that literature in a later section.
+
+\textbf{Compare to Choquet theory.} There is a straight forward link between our non-additive measure and Choquet theory.
+Recall that Choquet theory start by defining measure on the whole ensemble space before finding those constrained to the extreme points.
+Our setup is similar.
+The fraction capacity is defined on the whole ensemble space.
+In the classical discrete case, for example, if $A \subset \Omega \subset \Ens$ is an event, then $\frcap_{\ens}(A)$ will return the probability of event $A$; if $P \subset \Ens$ represents all the Poisson distributions with integer coefficient, $\frcap_{\ens}(P)$ will return how much of $\ens$ can be expressed as a mixture of Poisson distributions.
+In case that extreme points exist, or we can extend the ensemble space with some limit procedure, we can also define the fraction capacity on the extreme points.
 
 We can establish a direct connection between the fraction capacity and Choquet theory, because the ability to express an ensemble $\ens$ as the convex combination of $A$ is precisely the ability to find a measure $\mu$ that has support within $A$.
-That is, $\ens = \sum_{i=1}^{\infty} p_i \ens[a]_i$ if an only if the probability measure defined by $\mu(\{\ens[a]_i\}) = p_i$ represents $\ens$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{\mu(A) \, | \, \mu \in M_{\ens}\})$.\footnote{TODO: prove somewhere}
+That is, $\ens = \sum_{i=1}^{\infty} p_i \ens[a]_i$ if an only if the probability measure defined by $\mu(\{\ens[a]_i\}) = p_i$ represents $\ens$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{\mu(A) \, | \, \mu \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all possible measures.
 
@@
 This means that each element of the ensemble space is represented by a unique probability measure over the extreme points.
-In that case, the fraction capacity extended on the extreme points will also be additive.
+In that case, the fraction capacity restricted to the extreme points will also be additive.
 In the non-classical case, the closure of the ensemble space will not be a simplex, and there will be multiple representations, which will lead to a non-additive measure.
@@
 In other words, the fraction capacity reduces to a classical probability distribution over pure states in the classical case.
+
+\textbf{Uniqueness of the fraction capacity.} From a conceptual standpoint, the fraction capacity has a major advantage: it has a clear physical meaning and it is unique.
+Like in Choquet theory, and unlike for pseudo-probability, the domain over which the measure is defined is fully determined by the system, by the ensemble space.
+However, unlike Choquet, and like pseudo-probability once the space is fixed, there one unique fraction capacity for each ensemble.
+Moreover, the setup is more general than Choquet theory (i.e.
+the ensemble space need not be compact) and stems from necessary physical requirement.
+An in depth study on fraction capacities, then, is guaranteed to be applicable to all physical theories, including those yet to be formulated.
 

### 2025-07-10 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Added citations
Added references for the definition of Wigner, Kirkwood-Dirac, and SIC-POVVM representations.

Adhikary 2020 defines that:
"The state is non-classical
with respect to these observables if, even one of the
pseudo-probabilities in the scheme is negative."

That should be sufficient to support your claim that there is a focus on negative probabilities, but better references might exist. I started with Asthana 2021 and found references therein.

I properly added the citation of Hall for the entropic volume.

@@
 \textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions, of which the Wigner function is probably the most widely known.
-The Wigner transform $W : \Ens \to L^1(\mathbb{R}^2, \mathbb{R}) $, defined as [CITATION NEEDED]
+The Wigner transform $W : \Ens \to L^1(\mathbb{R}^2, \mathbb{R}) $, defined as \cite[eq.~11.8-3a]{mandel1995}
 \begin{equation}
@@
 In other circumstances, quasi-probability distributions over observables different from position and momentum are more useful.
-For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times.
+For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times \cite{lostaglio2023}.
 Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
 For example, the SIC-POVM represents quantum states as affine combinations of pure states $P_i$ that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
-Each ensemble $\rho$ can then be represented as
+Each ensemble $\rho$ can then be represented as \cite[eqs.
+28-29]{fuchs2004}
 \begin{equation}
@@
 \end{equation}
-where $p_i = \tr \left(P_i \rho\right)$. [CITATION NEEDED]
+where $p_i = \tr \left(P_i \rho\right)$.
 
@@
 
-\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasi-probability.
-[CITATION NEEDED] Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
+\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasi-probability \cite{adhikary2020}.
+Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
 This is a mischaracterization.
@@
 
-We now want to introduce a measure theoretic representation of the entropy that generalizes the notion of count of states.
+\textbf{Counting states.} We now want to introduce a measure theoretic representation of the entropy that generalizes the notion of count of states.
 Since entropy characterizes the variability of the preparations of the ensemble, greater variability means the ensemble is spread over more cases.
 Therefore the count of distinguishable states must be a monotonic function of the entropy.
-It has been shown[cite Hall] under very general conditions that the exponential of the entropy is the only function to have the appropriate features.
-One required feature, that we prove in general, is that the exponential of the entropy is subadditive (i.e.
-the count of states associated to a mixture cannot exceed the combined count of states associated to both mixtures) and it is additive for orthogonal ensembles (i.e.
-no overlap in possible instances) in a specific linear combination.
+Given that in statistical mechanics the entropy is the logarithm of the count of states, we expect the count of states to be the exponential of the entropy.
+Therefore one can argue that the proper way to measure the spread of an ensemble is through the exponential of the entropy, also known as entropic volume.\footnote{It has already been shown in \cite{hall1999} that the exponential of the entropy is the only function that can be used to define a volume.} However, the count of states, like the probability measure, is typically defined on the sample space $\Omega$, which can only work in the classical case.
+The entropy volume is a function of the ensemble space, so it is not a measure.
+Let's see everything fits together.
 
-Given a Borel set of ensembles $A \subseteq \Ens$, we define the \textbf{state capacity} of $A$ to be
+To build an intuition, consider the classical case.
+For every a region $U$ of phase space, we have a set of possible probability distributions that have support $U$.
+This is a convex set since every mixture of distributions with support $U$ has support $U$.
+Note that the uniform distribution $\rho_U$, if it exists, is the one with the greatest entropy, which will correspond to $S(\rho_U) = \log \mu(U)$.
+If $U$ is not compact, then there is no upper limit to the entropy and $\mu(U) = +\infty$.
+Now, let us transfer this intuition on the ensemble space $\ens$.
+If we take a set $A$, we can ask: what is the highest entropy reachable by a mixture of elements of $A$?
+If $A$ happens to be the set of ensembles with support $U$, then the supremum of the entropy is $\log \mu(U)$, which works also in the unbounded case.
+But the setup now is more general: $A$ can be any set of ensembles in any ensemble space.
+
+We still need to make sure that this definition still makes sense physically.
+If the entropic volume measure the spread of an ensemble, during a mixture the spread can at most sum.
+Remarkably, the bounds on the entropy are such that the exponential of the entropy is subadditive.
+That is, $2^{S(p\ens[a]+\bar{p}\ens[b])} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])}$ for any ensemble space.\footnote{It is an open question whether the exponential is the only monotonic function of the entropy that is subadditive.} Therefore this is indeed a general way to define the spread of ensembles and therefore the count of distinguishable states.
+
+\textbf{State capacity.} Given a Borel set of ensembles $A \subseteq \Ens$, we define the \textbf{state capacity} of $A$ to be
 \begin{equation}
@@
 \end{equation}
-This returns the exponential of the biggest entropy achievable with a mixture of elements of $A$.
+This returns the upper bound to the exponential of entropy achievable with a mixture of elements of $A$.
 
-One can then show that the fraction capacity $\capacity_{\ens} : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
+One can then show that the fraction capacity $\capacity : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
 \begin{enumerate}
@@
 In the discrete case, every point is associated to zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
-In the continuous case, gives a subset of phase space $A$, the maximum entropy is reached by the uniform distribution $\rho_A$, which means we recover $\log_2 \capacity(A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
+In the continuous case, gives a subset of phase space $U$ and $A_U$ the set of distributions with support within $U$, the maximum entropy is reached by the uniform distribution $\rho_U \in A_U$, which means we recover $\log_2 \capacity(A_U) = S(\rho_U)$, the fundamental postulate of statistical mechanics.
 Similarly, in quantum mechanics, given an orthogonal subspace $A$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(A) = 2^{\dim(A)}$.
@@
 The state capacity, then, recovers the number of distinguishable states.
-In classical mechanics, all states are perfectly distinguishable.
-In quantum mechanics, only orthogonal state are distinguishable, and therefore the state capacity recovers the number of orthogonal states.
+In classical mechanics, all states are perfectly distinguishable, the measure is additive.
+In quantum mechanics, only orthogonal state are distinguishable, therefore the measure is non-additive over pure states.
+We see again how the lack of single decomposition of ensembles over pure states is responsible for the failure of the additivity of the measure.
 

### 2025-07-10 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveMeasure: reviewed state capacity

@@
 \textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions, of which the Wigner function is probably the most widely known.
-The Wigner transform $W : \Ens \to L^1(\mathbb{R}^2, \mathbb{R}) $, defined as \cite[eq.~11.8-3a]{mandel1995}
+The Wigner transform $W : \Ens \to L^1(\mathbb{R}^2, \mathbb{R}) $, defined as [CITATION NEEDED]
 \begin{equation}
@@
 In other circumstances, quasi-probability distributions over observables different from position and momentum are more useful.
-For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times \cite{lostaglio2023}.
+For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times.
 Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
 For example, the SIC-POVM represents quantum states as affine combinations of pure states $P_i$ that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
-Each ensemble $\rho$ can then be represented as \cite[eqs.
-28-29]{fuchs2004}
+Each ensemble $\rho$ can then be represented as
 \begin{equation}
@@
 \end{equation}
-where $p_i = \tr \left(P_i \rho\right)$.
+where $p_i = \tr \left(P_i \rho\right)$. [CITATION NEEDED]
 
@@
 
-\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasi-probability \cite{adhikary2020}.
-Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
+\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasi-probability.
+[CITATION NEEDED] Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
 This is a mischaracterization.
@@
 Given that in statistical mechanics the entropy is the logarithm of the count of states, we expect the count of states to be the exponential of the entropy.
-Therefore one can argue that the proper way to measure the spread of an ensemble is through the exponential of the entropy, also known as entropic volume.\footnote{It has already been shown in \cite{hall1999} that the exponential of the entropy is the only function that can be used to define a volume.} However, the count of states, like the probability measure, is typically defined on the sample space $\Omega$, which can only work in the classical case.
+Therefore one can argue that the proper way to measure the spread of an ensemble is through the exponential of the entropy, also known as entropic volume.\footnote{It has already been shown[cite Hall] that the exponential of the entropy is the only function that can be used to define a volume.} However, the count of states, like the probability measure, is typically defined on the sample space $\Omega$, which can only work in the classical case.
 The entropy volume is a function of the ensemble space, so it is not a measure.

### 2025-07-10 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: started the section on non-additive theory

--- 2025-07-10 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-\section{Connection to fuzzy logic}
+\section{Non-additive generalization of probability theory}
 
-\section{Possible generalization of calculus}
+\textbf{Requirements.} We have shown how fraction capacities and the state capacities give us generalized version of probability measure and state counting measures.
+The question now becomes: is there a generalization of all other elements of probability theory (e.g.
+probability densities, random variables, expectation) to the non-additive case?
+Some of these elements require some form of calculus.
+Is the a non-additive version of calculus that would allow such a generalization?
+In this section we will describe what the ideal scenario would be and what are the problems that need to be addressed.
 
-As we pointed out, the state capacity and the fraction capacities are always well-defined, even in the cases where, potentially, Choquet theory does not apply because the closure of the ensemble space is not compact.
+First of all, we would need a guarantee that the capacities can always be restricted to the extreme points, and that the restriction is unique.
+As we say, the capacities are defined on the ensemble space $\Ens$, while the standard measures are defined on the space $X$ that supports the measures.
+As we said before, the ensemble space may not have extreme points: points of phase space do not correspond to a valid ensemble.
+We have been trying to understand how to reconstruct the points and the topology of $X$ from the space of absolutely continuous probability measures, but haven't been successful so far.\footnote{An attempt, for example, is by using orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
+However, this fails to recover the points correctly on the continuum even for classical mechanics.}
 
-General problems:
-1) Can we always find extreme points?
-2) Can we generalize to the non-compact case? Can \href{https://en.wikipedia.org/wiki/Krein%E2%80%93Milman_theorem#convex_compactness}{convex compactness} help?
-3) In the classical case, both measures become additive on the extreme points. Standard calculus is used. Is there a non-additive version of calculus that generalizes?
+Assuming we have a way to associate $\Ens$ with a set $X$ of extreme points (or equivalent notion), we can now think of the state capacity and the fraction capacity as non-additive measure over $X$.
+We have a notion of statistical variable $F : \Ens \to \mathbb{R}$ defined over the ensemble space, but in probability theory random variables are defined over the sample space.
+On the continuum, the expectation is recover through the integration over the probability density.
+Therefore, we are looking for a non-additive derivative $\rho = \frac{d \frcap_{\ens}}{d \capacity}$ that reduces to the Radon-Nikodym derivative in the additive case and non-additive integral such that
+\begin{equation}
+	F(\ens) = \int_X F(x) \rho(x) d\capacity.
+\end{equation}
+Though not a strict requirement, it would make sense if we found
+\begin{equation}
+	S(\ens) = \int_X \rho(x) \log \rho(x) d\capacity.
+\end{equation}
+
+Since the general problem is too hard to approach directly, a first step is to see whether such problem can be solved for the two-dimensional quantum case, the Bloch ball.
+There we can compute both the state capacity and the fraction capacity for different sets, and see whether some insights can be gained.
+Since there is already a rich literature on non-additive measures, we can see what results, if any, can be useful to our case.
 

### 2025-07-11 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Replace Wigner example with Husimi Q
The Husimi Q functions give a better example of functions where orthogonality doesn't imply disjoint support.

I added a reference to support our claim.

@@
 
-For example, the eigenstates $| m \rangle$ of the harmonic oscillator are orthogonal, mutually exclusive events. Their corresponding Wigner functions are
+For example, the Husimi Q functions of pure states $Q_\psi(\alpha) = \frac{1}{\pi} |\langle \alpha | \psi \rangle|^2$ have support everywhere, since we can express the inner product as \cite[eq.
+11.7-1]{mandel1995}
 \begin{equation}
-	W_m (q, p) = \frac{(-1)^m}{\pi} e^{-(q^2+p^2)} L_m (2(q^2 + p^2)).
+	\langle \alpha | \psi \rangle = e^{-|\alpha|^2/2} F_\psi(\alpha^\ast)
 \end{equation}
-Their support is the whole $\mathbb{R}^2$ since the Laguerre polynomials $L_m$ have finitely many zeros, and therefore are not disjoint.
+where $F_\psi$ is an entire analytic function.
 Conversely, in a SIC-POVM we take $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot be all orthogonal to each other.

### 2025-07-11 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reworded Husimi Q example

@@
 
-For example, the Husimi Q functions of pure states $Q_\psi(\alpha) = \frac{1}{\pi} |\langle \alpha | \psi \rangle|^2$ have support everywhere, since we can express the inner product as \cite[eq.
+For example, given a pure state $|\psi\>$, the Husimi Q function representing it is given by $Q_\psi(\alpha) = \frac{1}{\pi} |\langle \alpha | \psi \rangle|^2$.
+The inner product is given by \cite[eq.
 11.7-1]{mandel1995}
@@
 \end{equation}
-where $F_\psi$ is an entire analytic function.
-Conversely, in a SIC-POVM we take $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot be all orthogonal to each other.
+where $F_\psi$ is an entire analytic function, which cannot vanish on an open set without vanishing everywhere, therefore the support of all Husimi Q functions is the whole space $Z$.
+This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support; moreover disjoint subsets of $Z$ cannot be thought of as mutually exclusive events because event limited to any proper subset exists.
+For the discrete case, a SIC-POVM representation picks $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot be all orthogonal to each other.
 For a Bloch sphere, the four corners of the tetrahedron are not all opposite to each other.
 Therefore, disjoint sets of $P_i$ do not correspond to mutually exclusive events.
+Pseudo-probability representation, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
 

### 2025-07-14 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: adding proof that Choquet integral does not work

@@
 
+\textbf{Use of non-additive measure theory.} Given that extensive literature already exists for non-additive measure theory, we originally thought it would be just a matter of clarifying our definitions and connect them to the existing literature.
+However, things turned out to be more complicated, and we are now convinced that the tools already available are not applicable to our case.
+
+In this section we will first clarify the requirements and what we expect to have from a non-additive generalization of integration and derivation.
+Then we will show that the tools we have found in the literature, like the Sugeno and Choquet integral, fail for simple examples.
+
 \textbf{Requirements.} We have shown how fraction capacities and the state capacities give us generalized version of probability measure and state counting measures.
@@
 probability densities, random variables, expectation) to the non-additive case?
-Some of these elements require some form of calculus.
-Is the a non-additive version of calculus that would allow such a generalization?
-In this section we will describe what the ideal scenario would be and what are the problems that need to be addressed.
+Since integral and differential calculus in probability theory is based on measure theory, is the a non-additive version of calculus that would allow such a generalization?
 
-First of all, we would need a guarantee that the capacities can always be restricted to the extreme points, and that the restriction is unique.
-As we say, the capacities are defined on the ensemble space $\Ens$, while the standard measures are defined on the space $X$ that supports the measures.
-As we said before, the ensemble space may not have extreme points: points of phase space do not correspond to a valid ensemble.
-We have been trying to understand how to reconstruct the points and the topology of $X$ from the space of absolutely continuous probability measures, but haven't been successful so far.\footnote{An attempt, for example, is by using orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
-However, this fails to recover the points correctly on the continuum even for classical mechanics.}
+For simplicity, we will assume that Choquet theory applies: the ensemble space $\Ens$ has a clear set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
+In principle, the general problem is more complicated since, as we mentioned before, $\Ens$ may not have extreme points.
+Whether one can always reconstruct the points and the topology of $X$ given $\Ens$ is an open question that we leave for our larger work.\footnote{One attempt, for example, was by using orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
+However, this fails to recover the points correctly on the continuum even for classical mechanics.
+We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} For our purposes, we can assume that both the state capacity and the fraction capacity as non-additive measure over $X$.
 
-Assuming we have a way to associate $\Ens$ with a set $X$ of extreme points (or equivalent notion), we can now think of the state capacity and the fraction capacity as non-additive measure over $X$.
-We have a notion of statistical variable $F : \Ens \to \mathbb{R}$ defined over the ensemble space, but in probability theory random variables are defined over the sample space.
-On the continuum, the expectation is recover through the integration over the probability density.
-Therefore, we are looking for a non-additive derivative $\rho = \frac{d \frcap_{\ens}}{d \capacity}$ that reduces to the Radon-Nikodym derivative in the additive case and non-additive integral such that
+We goal is to define an expectation for statistical variables that reduces to the standard case in classical theories, that is, in the additive case.
+We already have a notion of statistical variable $F : \Ens \to \mathbb{R}$ defined over the ensemble space.
+This too can be restricted on the extreme points $X$ and, given that $F$ is a linear functional, the values on the extreme points will fully determine the values for all other points of $\Ens$.
+In the classical case, the expectation is recovered through the integration of the variable over the probability density.
+Therefore, we would like a non-additive integral such that
 \begin{equation}
-	F(\ens) = \int_X F(x) \rho(x) d\capacity.
+	E[F] = F(\ens) = (E)\int_X F(x) d\frcap_{\ens}.
 \end{equation}
-Though not a strict requirement, it would make sense if we found
+which reduces to the Riemann integral in the classical (additive) case.
+The integral will depend only on the values of $F$ and $\frcap_{\ens}$ on $X$, but since these uniquely determine both $F$ and $\ens$, the integral can be understood as a map from $F$ and $\ens$ to $F(\ens)$, which is, in principle, well-defined.
+Therefore, in principle, this integral should exist.
+
+Additionally, we would like to be able to define an analogue of the probability density.
+We are therefore looking for a non-additive derivative $\rho = \frac{d \frcap_{\ens}}{d \capacity}$ that reduces to the Radon-Nikodym derivative in the classical (additive) case such that
 \begin{equation}
-	S(\ens) = \int_X \rho(x) \log \rho(x) d\capacity.
+	F(\ens) = (E)\int_X F(x) \rho(x) d\capacity
+\end{equation}
+under the previous integral.
+It is less clear whether this derivative is always well defined.
+Since in classical mechanics we have $S(\ens) = \int_X \rho \log \rho d\mu$ and in quantum mechanics $S(\ens) = \tr[\rho \log \rho]$, we would expect
+\begin{equation}
+	S(\ens) = (E)\int_X \rho(x) \log \rho(x) d\capacity.
+\end{equation}
+We do not know whether this second requirement is at all possible, or whether it can be used as guide in case that the previous derivative is not uniquely defined.
+
+\textbf{Quick overview of the literature.}
+
+\textbf{A test case for the Choquet integral.} We are not going to compute the Choquet integral for a two state quantum system and see that, unfortunately, it does not recover the expectation of the statistical variables in the corresponding ensemble space.
+
+The ensemble space for a two-state quantum system is the Bloch ball.
+The extreme points form a sphere and all other ensembles are the internal points of the sphere.
+The center point $\ens[o]$ is the maximally mixed state.
+Since $\ens[o]$ can always be obtained by the equal mixture of two opposite points, we have that the fraction capacity for every singleton of an extreme point is $\frcap_{\ens[o]}(\{\psi\}) =\frac{1}{2}$.
+A statistical variable $F$, instead is a linear function of the space.
+This means that two opposite points on the surface $\ens[a]$ and $\ens[b]$ will take the minimum and maximum values $a = F(\ens[a])$ and $b = F(\ens[b])$.
+For simplicity, we can picture $\ens[a]$ and $\ens[b]$ as the extreme points of the vertical axis.
+
+The Choque integral is defined as 
+\begin{equation}
+	\begin{aligned}
+(C)&\int F d\frcap_{\ens[o]} := \\
+&\int_{-\infty}^0
+(\frcap_{\ens[o]} (\{x | F (x) \geq s\})-\frcap_{\ens[o]}(X))\, ds \\
+&+
+\int^\infty_0
+\frcap_{\ens[o]} (\{x | F (x) \geq s\})\, ds.
+	\end{aligned}
+\end{equation}
+The set $\{x | F (x) \geq s\}$ corresponds to all the points of $X$, the surface of the sphere, that have a value of $F$ greater than a given $s$.
+Since $F$ is a linear function, this corresponds to the set of points above a given height.
+So, all we need to do to calculate the integral is calculate the fraction capacity for sets of sphere $A(s)$ where the bottom is cut off at all possible values.
+
+If $s\leq \frac{1}{2} F(\ens[a]) + \frac{1}{2} F(\ens[b])$, we are cutting the sphere at the midpoint or below.
+This means that $A(s)$ will include two opposite points.
+Since $\ens[o]$ can be expressed as the mixture of two opposite point, $\ens[o]$ is in the hull of $A(s)$ and therefore $\frcap_{\ens[o]}(A(s)) = 1$.
+Conversely, if $s > F(\ens[b])$, $A(s) = \emptyset$, which means $\frcap_{\ens[o]} = 0$.
+The non-trivial case, then, is when $\frac{1}{2} F(\ens[a]) + \frac{1}{2} F(\ens[b]) < s \leq F(\ens[b])$.
+
+In this case, the hull of $A(s)$ is the top spherical cap.
+We need to consider all possible convex combinations $\ens[o] = p \ens + \bar{p} \ens[c]$ where $\ens$ is in the spherical cap and $\ens[c]$ is not, and find the one with highest $p$.
+Since, for combinations along the same direction, $p$ increases as $\ens$ gets closer to $\ens[o]$ and as $\ens[c]$ gets further, $\ens[c]$ will be an extreme point on the lower hemisphere while $\ens$ will be on the disc at constant $s$, the lower side of the cap.
+Also note that the distance between $\ens[o]$ is the same for all extreme points, therefore, to maximize $p$, we simply need to minimize the distance between $\ens[o]$ and $\ens$.
+We have that $\ens$ must be the intersection between the vertical axis and the disk, and $\ens[c]$ will coincide with $\ens[a]$.
+The coefficient $p$ will be given by ratio of the distance $\overline{\ens[o]\ens[a]}$ and $\overline{\ens\ens[a]}$.
+Since $F$ is a linearly functional, we have
+\begin{equation}
+	\frac{\overline{\ens[o]\ens[a]}}{\overline{\ens\ens[a]}} = \frac{F(\ens[o])-F(\ens[a])}{F(\ens) - F(\ens[a])} = \frac{1}{2} \frac{F(\ens[b]) - F(\ens[a])}{s-F(\ens[a])}.
 \end{equation}
 
-Since the general problem is too hard to approach directly, a first step is to see whether such problem can be solved for the two-dimensional quantum case, the Bloch ball.
-There we can compute both the state capacity and the fraction capacity for different sets, and see whether some insights can be gained.
-Since there is already a rich literature on non-additive measures, we can see what results, if any, can be useful to our case.
+To recap, we have
+\begin{equation}
+	\frcap_{\ens[o]}(A(s))=
+	\begin{cases}
+		1 & \text{if } s\leq \frac{1}{2} a + \frac{1}{2} b\\
+		\frac{1}{2} \frac{b - a}{s-a} & \text{if } \frac{1}{2} a + \frac{1}{2} b < s \leq b\\
+		0 & \text{if } s > b.
+	\end{cases}
+\end{equation}
+We can now calculate the integral
+\begin{equation}
+	\begin{aligned}
+		(C)&\int F d\frcap_{\ens[o]} := 
+		\int_{-\infty}^0
+		(\frcap_{\ens[o]} (A(s))-\frcap_{\ens[o]}(X))\, ds \\
+		&+
+		\int^\infty_0
+		\frcap_{\ens[o]} (A(s))\, ds \\
+		&= \int_{-\infty}^0 (1 - 1)ds + \int_{0}^{\frac{1}{2} a + \frac{1}{2} b}  1 ds \\
+		&+ \int_{\frac{1}{2} a + \frac{1}{2} b}^{b}  \frac{1}{2} \frac{b - a}{s-a} ds + \int_{b}^{\infty} 0 ds \\
+		&=\left[s\right]_{0}^{\frac{1}{2} a + \frac{1}{2} b}
+		+ \frac{1}{2}\left(b - a\right)\left[\ln (s - a)\right]_{\frac{1}{2} a + \frac{1}{2} b}^{b} \\
+		&= \frac{1}{2} a + \frac{1}{2} b + \frac{1}{2}\left(b - a\right)\left[\ln (b - a)\right. \\
+		&+\left. \ln (\frac{1}{2} a + \frac{1}{2} b - a) \right] \\
+		&= \frac{1}{2} a + \frac{1}{2} b + \frac{1}{2}\left(b - a\right)\ln (2)
+	\end{aligned}
+\end{equation}
+
+Note that, since $\ens[o]$ is the center of the ball, $F(\ens[o]) = \frac{1}{2} a + \frac{1}{2} b \neq \frac{1}{2} a + \frac{1}{2} b + \frac{1}{2}\left(b - a\right)\ln (2)$.
+This means that the Choquet integral of the statistical variable with respect to the fraction capacity does not recover the expectation value.
 

### 2025-07-14 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: adding draft figures

@@
 \usepackage{tcolorbox}
+\usepackage{float}
 
@@
 
-\textbf{A test case for the Choquet integral.} We are not going to compute the Choquet integral for a two state quantum system and see that, unfortunately, it does not recover the expectation of the statistical variables in the corresponding ensemble space.
+\textbf{A test case for the Choquet integral.} We are now going to compute the Choquet integral for a two state quantum system and see that, unfortunately, it does not recover the expectation of the statistical variables in the corresponding ensemble space.
 
-The ensemble space for a two-state quantum system is the Bloch ball.
-The extreme points form a sphere and all other ensembles are the internal points of the sphere.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{ChoquetCalculation1.jpg}
+	\caption{...}
+\end{figure}
+
+The ensemble space $\Ens$ for a two-state quantum system is the Bloch ball.
+The pure states, the extreme points, are on the surface $X$ and the internal points correspond to mixed states.
 The center point $\ens[o]$ is the maximally mixed state.
-Since $\ens[o]$ can always be obtained by the equal mixture of two opposite points, we have that the fraction capacity for every singleton of an extreme point is $\frcap_{\ens[o]}(\{\psi\}) =\frac{1}{2}$.
-A statistical variable $F$, instead is a linear function of the space.
-This means that two opposite points on the surface $\ens[a]$ and $\ens[b]$ will take the minimum and maximum values $a = F(\ens[a])$ and $b = F(\ens[b])$.
+Since $\ens[o]$ can always be obtained by the equal mixture of two opposite points (i.e.
+two orthogonal states), the fraction capacity for every singleton of an extreme point $\psi$ is $\frcap_{\ens[o]}(\{\psi\}) =\frac{1}{2}$.
+A statistical variable $F$ (i.e.
+an observables), is a linear function of the ensemble space.
+This means that two opposite points on the surface $\ens[a]$ and $\ens[b]$ (i.e.
+the eigenstates) will take the minimum and maximum values $a = F(\ens[a])$ and $b = F(\ens[b])$.
 For simplicity, we can picture $\ens[a]$ and $\ens[b]$ as the extreme points of the vertical axis.
 
-The Choque integral is defined as 
+The Choquet integral is defined as 
 \begin{equation}
@@
 \end{equation}
-The set $\{x | F (x) \geq s\}$ corresponds to all the points of $X$, the surface of the sphere, that have a value of $F$ greater than a given $s$.
+The set $A(s) = \{x | F (x) \geq s\}$ corresponds to all the points of $X$, the surface of the sphere, that have a value of $F$ greater than a given $s$.
 Since $F$ is a linear function, this corresponds to the set of points above a given height.
@@
 
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{ChoquetCalculation2.jpg}
+	\caption{...}
+\end{figure}
+
 In this case, the hull of $A(s)$ is the top spherical cap.
@@
 \begin{equation}
-	\frac{\overline{\ens[o]\ens[a]}}{\overline{\ens\ens[a]}} = \frac{F(\ens[o])-F(\ens[a])}{F(\ens) - F(\ens[a])} = \frac{1}{2} \frac{F(\ens[b]) - F(\ens[a])}{s-F(\ens[a])}.
+	\frac{\overline{\ens[o]\ens[a]}}{\overline{\ens\ens[a]}} = \frac{F(\ens[o])-F(\ens[a])}{F(\ens) - F(\ens[a])} = \frac{1}{2} \frac{b - a}{s-a}.
 \end{equation}
@@
 		&= \frac{1}{2} a + \frac{1}{2} b + \frac{1}{2}\left(b - a\right)\left[\ln (b - a)\right. \\
-		&+\left. \ln (\frac{1}{2} a + \frac{1}{2} b - a) \right] \\
+		&+\left. \ln \left(\frac{1}{2} a + \frac{1}{2} b - a\right) \right] \\
 		&= \frac{1}{2} a + \frac{1}{2} b + \frac{1}{2}\left(b - a\right)\ln (2)

### 2025-07-15 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: adding some literature for non-additive measures

@@
 
+One important feature, for example, would be to characterize what fraction capacities are possible for each ensembles space.
+In the classical case, for example, only additive fraction capacities are possible; in the quantum case, only some type of non-additivity will be possible.
+The shape of the space effectively is encoded in the space of possible fraction capacities, maybe in a way that is mathematically and physically insightful.
+
 \section{State capacity}
@@
 
-\textbf{Quick overview of the literature.}
+\textbf{Ruling out Sugeno and Choquet integrals.} Non-additive measures are used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence.
+CITE To our knowledge, there isn't a single unified theory, though some work has been proceeding in that direction.\cite{nonadditive2014,pap2013} For now, we will concentrate on integration.
+The Choquet integral and the Sugeno integral are the oldest and most common approaches to integration of non-additive measure, and therefore we should first test whether these work.
 
-\textbf{A test case for the Choquet integral.} We are now going to compute the Choquet integral for a two state quantum system and see that, unfortunately, it does not recover the expectation of the statistical variables in the corresponding ensemble space.
+Note that the Sugeno integral does not recover the Lebesgue integral in the case the capacity is additive (i.e.
+classical), and is therefore ruled out.
+The Choquet integral does, so it will work in the classical case.
+However, it does not work in the quantum case.
+We can show this by computing the Choquet integral for a two states quantum system and see that it does not recover the expectation of the statistical variables in the corresponding ensemble space.
 

### 2025-07-16 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reworked introduction to non-additive integrals

@@
 \end{equation}
+
+\textbf{Expectation and integration.} TODO: show the use of integrals over probability measure.
 
@@
 
-\textbf{Use of non-additive measure theory.} Given that extensive literature already exists for non-additive measure theory, we originally thought it would be just a matter of clarifying our definitions and connect them to the existing literature.
-However, things turned out to be more complicated, and we are now convinced that the tools already available are not applicable to our case.
+\textbf{Requirements.} In a previous section we reviewed what types of objects we have in probability theory for classical ensembles: a sample space $X$, a probability measure $p$, random variables $F : X \to \mathbb{R}$, expectations as integrals of the probability measure $E[F] = \int_X F(x) dp$, a state counting measure $\mu$, the probability density as the Radon-Nikodym derivative $\frac{dp}{d\mu}$ and the entropy as $\int_X \rho \log \rho d\mu$.
+Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the states counting measure $\mu$, is there a generalization of all other elements?
+That is, can we find an integral and derivative operation over non-additive measures that generalizes to all physical theories?
 
-In this section we will first clarify the requirements and what we expect to have from a non-additive generalization of integration and derivation.
-Then we will show that the tools we have found in the literature, like the Sugeno and Choquet integral, fail for simple examples.
+The first difficulty, as we already pointed out, is that the ensemble space $\Ens$ may not have extreme points.
+What is the correct procedure to ``close'' the ensembles space and recover the space of extreme points is, for us, an open question.
+\footnote{One attempt, for example, was by using orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
+However, this fails to recover the points correctly on the continuum even for classical mechanics.
+We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} For the purpose of this paper, we will assume $\Ens$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
+Therefore we can talk about the restriction on $X$ of the state capacity, the fraction capacity and the statistical variables.
 
-\textbf{Requirements.} We have shown how fraction capacities and the state capacities give us generalized version of probability measure and state counting measures.
-The question now becomes: is there a generalization of all other elements of probability theory (e.g.
-probability densities, random variables, expectation) to the non-additive case?
-Since integral and differential calculus in probability theory is based on measure theory, is the a non-additive version of calculus that would allow such a generalization?
-
-For simplicity, we will assume that Choquet theory applies: the ensemble space $\Ens$ has a clear set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
-In principle, the general problem is more complicated since, as we mentioned before, $\Ens$ may not have extreme points.
-Whether one can always reconstruct the points and the topology of $X$ given $\Ens$ is an open question that we leave for our larger work.\footnote{One attempt, for example, was by using orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
-However, this fails to recover the points correctly on the continuum even for classical mechanics.
-We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} For our purposes, we can assume that both the state capacity and the fraction capacity as non-additive measure over $X$.
-
-We goal is to define an expectation for statistical variables that reduces to the standard case in classical theories, that is, in the additive case.
-We already have a notion of statistical variable $F : \Ens \to \mathbb{R}$ defined over the ensemble space.
-This too can be restricted on the extreme points $X$ and, given that $F$ is a linear functional, the values on the extreme points will fully determine the values for all other points of $\Ens$.
-In the classical case, the expectation is recovered through the integration of the variable over the probability density.
-Therefore, we would like a non-additive integral such that
+Recall that a statistical variable is linear function $F : \Ens \to \mathbb{R}$ defined over the ensemble space.
+Given that $F$ is a linear, its restriction on $X$ will determine the values for all other points of $\Ens$.
+Since the expectation is recovered through the integration of the variable over the probability density, we would like a non-additive integral $(E)\int_X$ such that
 \begin{equation}
@@
 \end{equation}
-under the previous integral.
-It is less clear whether this derivative is always well defined.
-Since in classical mechanics we have $S(\ens) = \int_X \rho \log \rho d\mu$ and in quantum mechanics $S(\ens) = \tr[\rho \log \rho]$, we would expect
+under the previous integral. It is less clear whether this derivative is always well defined. Ideally, we would also expect
 \begin{equation}
@@
 \end{equation}
-We do not know whether this second requirement is at all possible, or whether it can be used as guide in case that the previous derivative is not uniquely defined.
+This last requirement may not be possible, or it may be used as guide in case that the previous derivative is not uniquely defined.
 
-\textbf{Ruling out Sugeno and Choquet integrals.} Non-additive measures are used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence.
-CITE To our knowledge, there isn't a single unified theory, though some work has been proceeding in that direction.\cite{nonadditive2014,pap2013} For now, we will concentrate on integration.
-The Choquet integral and the Sugeno integral are the oldest and most common approaches to integration of non-additive measure, and therefore we should first test whether these work.
+\textbf{Ruling out Sugeno and Choquet integrals.} Given that extensive literature already exists for non-additive measure theory (cite), as it is used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence, we originally thought it would be just a matter of clarifying our definitions and connect them to the existing literature.
+However, things turned out to be more complicated.
+Depending on the use cases, different types of integrals exists.
+In fact, procedures exist\cite{nonadditive2014,pap2013} to construct infinitely many types of non-additive integrals.\footnote{Briefly, one can define a pseudo-addition and a pseudo-multiplication as real functions of two real arguments and construct an integral based on those definitions.} We can, however, at least rule out the two most common non-additive integrals: the Sugeno integral and the Choquet integral.
 
-Note that the Sugeno integral does not recover the Lebesgue integral in the case the capacity is additive (i.e.
-classical), and is therefore ruled out.
+The Sugeno integral does not recover the Lebesgue integral in the case the capacity is additive (i.e.
+classical), and is therefore quickly ruled out.
 The Choquet integral does, so it will work in the classical case.
@@
 
+\textbf{Ruling out Sugeno and Choquet integrals.}
+
 \section{Conclusion}

### 2025-07-17 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability

@@
 
-\textbf{Ensemble space embedding and signed measures.} We have already seen that we can represent ensembles using classical probability if and only the ensemble space is a simplex.
-Since any probability measure is possible, we can assign probability to two disjoint areas independently.
-This answers the first question.
+\textbf{Ensemble space embedding and signed measures.} As we saw when reviewing Choquet theory, and measure on the extreme points represents an ensemble.
+Classical spaces are those in which there is a one-to-one correspondence between each measure and each ensemble.
+We also saw that the hallmark of pseudo-probabilities is the lack of mutual exclusivity of the events.
+Why is mutual exclusivity linked to the lack of multiple representations?
+
+Suppose we have an ensemble $\ens$ that can be represented as different measures $p_1$ and $p_2$ over the extreme points.
+This means that there are two sets $A$ and $B$ for which $p_1(A) < p_2(A)$ and $p_1(B) > p_2(B)$.
+That is, we can move probability from $A$ to $B$ while leaving the ensemble, and therefore the entropy and all statistical variables, unchanged.
+If $A$ and $B$ where two mutually exclusive events, this would not be possible: if we could distinguish between $A$ and $B$, something would have to change.
+The only way for $A$ and $B$ to be mutually exclusive, then, is if different measures represent different ensembles, which means the ensemble space is a simplex.
+This answers the first question: we can use classical probability exactly when all extreme points are mutually exclusive (i.e.
+orthogonal as defined from the entropy) and the ensemble space is a simplex.
+
+In quantum mechanics, a classical context is a simplex embedded within the ensemble space.
+For the Bloch ball, a classical context is an axis that connect two opposite point.
+The points are orthogonal because the midpoint is the maximally mixed state, which therefore maximizes the entropy increase, and the line segment is a one simplex.
+If we take the position operator and consider all density operators that commute with $x$, this is exactly functions $\rho(x)$ that integrate two once, which is the space of probability supported by $X$, which is a simplex.
+We want to stress how the conceptual view and the mathematical view fits precisely together with little effort.
 
@@
 
-To recap, we have
+To recap, we have (TODO: add a graph)
 \begin{equation}
@@
 
-\textbf{Ruling out Sugeno and Choquet integrals.}
+\textbf{Space of capacities and variables.} A bigger conceptual problem is that the space of possible fraction capacities of an ensemble space depends on the ensemble space.
+For a classical ensemble space, for example, only additive measures are allowed.
+Conversely, all statistical variables are linear on the ensemble space, so the restriction on the extreme points cannot be arbitrary.
+In the Bloch ball case, for example, all linear functions are fully determined by two opposite points on the sphere and by two value (i.e.
+the eigenstates and the eigenvalues of the corresponding operator).
+The value for all the other points is determined.
+
+This again corresponds to assumptions of mutual exclusivity of the extreme points and the corresponding events.
+As we saw before, the additivity of the measure corresponds to the ability to assign probability to a set without implicitly assigning probability to another.
+In terms of statistical variable, two mutually exclusive events can be assigned values independently.
+This, again, recovers how observable work in quantum mechanics, without mentioning quantum mechanics: only orthogonal states, that is mutually exclusive events, can associated to measurement outcomes and be given a classical probability distribution.
+The non-additivity of the fraction capacity and the non-arbitrariness of the statistical variables over the extreme points, then, must be linked to the shape of the ensemble space.
+To our knowledge, there is no such requirement for the more typical used of non-additive measure theory.
+
+Additionally, the non-additivity of the state capacity also depends on the degree of mutually exclusiveness of the events.
+It is additive for mutually exclusive events and sub-additive otherwise.
+Therefore state capacity, fraction capacity and statistical variables have to be, in a sense, compatible with each other.
+
+As we saw in the classical case, probability measures have to be absolutely continuous with respect to the state counting measure.
+Also, if the state capacity is additive, all disjoint events are mutually exclusive, which means the fraction capacity is additive and disjoint events can correspond to independent values for a statistical variable.
+Therefore, we wonder whether a more general compatibility condition exists between the fraction capacity and the state capacity/statistical variable respectively.
+
+We have not found similar discussion in the literature for non-additive measure, which makes us suspect that these considerations are specific to this setting.
 

### 2025-07-18 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed ensemble space

@@
 
-We start by assuming that physical laws deal with experimentally reproducible relationships, which means that they must describe relationships between statistical ensembles.
+We start by assuming that physical laws deal with experimentally reproducible relationships, which implicitly requires the notion of statistical ensembles.
 Conceptually, a statistical ensemble is the collection of outputs, taken at once, of a reproducible preparation procedure for a physical system.
@@
 a positive semi-definite self-adjoint operator of trace one).
-An ensemble space is the collection of all possible ensembles defined in a physical theory, together with a basic mathematical structure that all physical theories must possess.
+A physical law, then, ultimately expresses relationship among ensembles.
+A minimal requirement for a physical theory, then, is to provide an ensemble space: the collection of all possible ensembles within the theory, together with a basic mathematical structure that all physical theories must possess.
 
-An ensemble space $\Ens$ will need to satisfy a three basic requirements.
+An ensemble space $\Ens$ will need to satisfy three basic requirements.
 First, it needs to guarantee experimental verifiability.
@@
 
-Additionally, we will require that ensembles can be identified through real valued statistical quantities.
-\footnote{It is an open question whether all ensemble spaces must possess this property.
-In a previous work we showed that the dense linear ordering require highly idealized assumptions.} A statistical quantity represents the expectation of a real valued quantity, and is therefore an affine continuous functions $F : \Ens \to \mathbb{R}$.
+Additionally, we will require that ensembles can be identified through real valued statistical quantities.\footnote{It is an open question whether this is an additional assumption or it can be derived from the other requirements.} A statistical quantity represents the expectation of a real valued quantity, and is therefore an affine continuous functions $F : \Ens \to \mathbb{R}$.
 That is, $F(p \ens[a] + \bar{p} \ens[b]) = p F(\ens[a]) + \bar{p} F(\ens[b])$.
 The last requirement is that there exists a family $F_i$ of statistical quantities that fully identify each ensemble.
-That is, $F_i(p \ens[a] + \bar{p} \ens[b]) = p F_i(\ens[a]) + \bar{p} F_i(\ens[b])$ and $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.
+That is, $F_i(p \ens[a] + \bar{p} \ens[b]) = p F_i(\ens[a]) + \bar{p} F_i(\ens[b])$ and $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.\footnote{For example, on the continuum this means that the moment problem is solvable: you can identify the distribution by the expectation of the polynomials of position and momentum.
+Similarly, in quantum mechanics means that the state can be reconstructed through quantum tomography.}
 
-These requirements interact with each other leading to a rich mathematical structure that connect to many of the structures we use in mathematical physics.
-In this paper, we will examine only connections to measure theoretic representations.
+Each physical theory satisfies these axioms, even though it may be represented very different mathematically.
+Classical mechanics, for example, is typically described in terms of the geometry of phase space.
+Quantum mechanics is most often defined in terms of complex inner product spaces, but can be reframed in terms of quasi-probability distributions.
+Therefore, while the above requirements are needed for a physical theory, it does not necessarily mean that it is convenient to describe problems or make calculations in terms of those definitions.
+In fact, those requirements interact with each other leading to a rich mathematical structure that connect to many of the structures we use in mathematical physics, and its goal is mainly to see how the different representations are connected.
 
-\textbf{Ensemble spaces as convex sets.} In this work, an ensemble space $\Ens$ is a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{The entropy constraints forces the topology to be $\mathsf{T}_1$ the convex space to embed into a vector space.
+The goal of this paper is to examine the different way that measure theory can be used to represent ensembles in any given physical theory.
+
+\textbf{Ensemble spaces as convex sets.} For the purpose of measure theoretic representations, it is best to understand an ensemble space $\Ens$ as a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{The entropy constraints forces the topology to be $\mathsf{T}_1$ the convex space to embed into a vector space.
 A second countable $\mathsf{T}_1$ TVS is metrizable.
-The family of statistical quantities induce a family of semi-norms on the space, which must be countable due to second countability, which makes the TVS locally convex.} The entropy is function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
+The family of statistical quantities induce a family of semi-norms on the space, which must be countable due to second countability, which makes the TVS locally convex.} The entropy is a function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
 \begin{equation}
@@
 
+Using the convex structure, we say that two ensembles are separate, noted $\ens[a] \separate \ens[b]$, if they cannot be expressed as a mixture of a common ensemble.
+That is, there is no $\ens[c]$ such that $\ens[a] = \alpha \ens[c] + \bar{\alpha} \ens[d]$ and $\ens[b] = \beta \ens[c] + \bar{\beta} \ens$ from some $\alpha, \beta \in (0,1]$ and $\ens[d],\ens \in \Ens$.
+No ensemble is separate from itself.
+An internal point is an ensemble that is not separate from any ensemble (i.e.
+it is the mixture of any ensemble).
+An extreme point is an ensemble that is separate from all others (i.e.
+it is only a mixture of itself).
+
+TODO: figure with classical triangle and quantum Bloch Ball
+
+The set of all extreme points $X$, as we will see, plays a major role.
+However, on physical grounds, these may not be proper ensembles.
+In classical mechanics, these would corresponds to Dirac measure: probability distributions all concentrated at a single point of phase space.
+Since we cannot reliably prepare a particle with infinite precision, these are not ensemble and they are not part of $\Ens$.
+For the purpose of Choquet theory, we will assume that there is a way to find the closure $E \supset \Ens$ that includes all extreme points $X$.
+Moreover, we will also assume $E$ to be compact, even though it will not be.
+How to properly make these identification, and how to go beyond the compact case, is outside of the scope of this particular article.\footnote{An early attempt used the notion of orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
+This does not work.
+An approach, currently under investigation, is to study limits over statistical quantities.
+We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} Here we will use $\Ens$ to indicate the general case and $E$ to indicate this ``compact closure.''
+
 \textbf{Orthogonality and separatedness.} We will say that two ensembles are orthogonal, noted $\ens[a] \ortho \ens[b]$, when their mixture gives a maximal entropy increase.
 This will recover the notion of orthogonality for both classical and quantum spaces.
-Another notion of distinction between two ensembles is in terms of the mixtures: two ensembles are separate, noted $\ens[a] \separate \ens[b]$, they cannot be expressed as a mixture of a common ensemble.
-That is, there is no $\ens[c]$ such that $\ens[a] = \alpha \ens[c] + \bar{\alpha} \ens[d]$ and $\ens[b] = \beta \ens[c] + \bar{\beta} \ens$ from some $\alpha, \beta \in (0,1]$ and $\ens[d],\ens \in \Ens$.
-In classical spaces, these two notions coincide: two ensembles are orthgonal and separate if and only if the support of the respective probability distributions overlap on a non-measure-zero set.
+Conceptually, these corresponds to two ensembles where all instances are distinct from the other: they are mutually exclusive and a single instance is enough to distinguish between the two cases.
+
+In classical spaces, orthogonality and separatedness coincide: two ensembles are orthgonal and separate if and only if the support of the respective probability distributions overlap on a non-measure-zero set.
 In quantum mechanics, these two notions are different: any two pure states are separate but are not orthogonal.
+For example, pure ensembles of spin up and spin down are orthogonal, mutually exclusive, precisely because it takes one instance to determine which one is which.
+Pure ensembles of spin up and spin right, however, are not orthogonal precisely because, in some cases, their instance cannot be perfectly discriminated.
+They are, however, orthogonal, because they are both pure states, one is not the mixture of other ensembles: they are extreme points.
+
+It is the ease with which such distinctions between the classical and quantum case can be made that is the main appeal for this structure.
+We will see over and over how mutual exclusivity is linked to classical probability and its failure in quantum, and other non-classical, theories.
 
@@
 Under these assumptions, an ensemble space $\Ens$ is a convex set of a metrizable second countable locally convex topological vector space $V$.
+In some cases, for simplicity, we will require $\Ens$ to admit a ``compact closure'' $E$, so that it includes the set $X$ of its extreme points.
 We now turn to the question of how we can represent both the entropy and each ensemble in terms of some measure theoretic structure.

### 2025-07-21 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reworked Choquet theory

@@
 
+% space is metrizable https://proofwiki.org/wiki/Birkhoff-Kakutani_Theorem/Topological_Vector_Space
+
 \textbf{Ensemble spaces as convex sets.} For the purpose of measure theoretic representations, it is best to understand an ensemble space $\Ens$ as a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{The entropy constraints forces the topology to be $\mathsf{T}_1$ the convex space to embed into a vector space.
@@
 In classical mechanics, these would corresponds to Dirac measure: probability distributions all concentrated at a single point of phase space.
-Since we cannot reliably prepare a particle with infinite precision, these are not ensemble and they are not part of $\Ens$.
+Since we cannot reliably prepare a particle with infinite precision, and its entropy would correspond to minus infinity, these are not ensemble and they are not part of $\Ens$.
 For the purpose of Choquet theory, we will assume that there is a way to find the closure $E \supset \Ens$ that includes all extreme points $X$.
-Moreover, we will also assume $E$ to be compact, even though it will not be.
+Moreover, we will also assume $E$ to be compact, even though it will not be in general.
 How to properly make these identification, and how to go beyond the compact case, is outside of the scope of this particular article.\footnote{An early attempt used the notion of orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
@@
 An approach, currently under investigation, is to study limits over statistical quantities.
-We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} Here we will use $\Ens$ to indicate the general case and $E$ to indicate this ``compact closure.''
+We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} Note that the classical discrete and the quantum case fall in the compact case.
+Moreover, the set of all probability distributions over a compact space $X$ also fall in the compact case.
+Therefore, apart from the technicality above, the compact case is still a very interesting case.
+To distinguish the context, in this article we will use $\Ens$ to indicate the general case and $E$ to indicate this ``compact closure.''
 
@@
 
-\textbf{Classical probability measures.} In the classical case, ensembles are defined as probability measures of the space of all possible states $X$.
+\textbf{Classical probability measures.} In the classical case, given the topological space of possible states $X$ (i.e.
+the sample space), ensembles are defined as probability measures $p : \Sigma_X \to [0,1]$.
+That is, a probability measure assigns to each element of the Borel algebra $\Sigma_X$\footnote{In our overall project, Borel sets represent statements associated to a test, regardless of termination.}, an event, the probability to find it to be true.
 In the discrete case, like for roll of a die, $X$ is a set endowed with the discrete topology\footnote{Physically, the discrete topology corresponds to the ability to experimentally verify or falsify each case.
-That is, we are able to recognize whether a particular die landed or didn't land on a particular face.} and, consequently, the power set as the $\sigma$-algebra.
-In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variable which are verifiable up to finite but arbitrarily small precision.
-That is, we are able to tell whether a state is within a particular range of position and momentum.} and the standard Borel sets as the $\sigma$-algebra.
-In either case, a probability measure is a map $p : \Sigma_X \to [0,1]$ that assign to each event of the $\sigma$-algebra (i.e.
-to an empirical statements) the probability to find it to be true.
-The probability measure is additive, meaning that if two events $A, B \in \Sigma_X$ are mutually exclusive (i.e.
-$A \cap B = \emptyset$) then the probability of the union is the sum of the probability of the respective events:
+That is, we are able to recognize whether a particular die landed or didn't land on a particular face.} and, consequently, the power set is the $\sigma$-algebra.
+In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold.\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variable which are verifiable up to finite but arbitrarily small precision.}  The probability measure is countably additive, meaning that, given a countable collection $A_i$ of pair-wise disjoint events (i.e.
+$A_i \cap A_j = \emptyset$), the probability of the union is the sum of the probability of the respective events:
 \begin{equation}
-	p(A\cup B) = p(A) + p(B).
+	p\left(\bigcup_i A_i \right) = \sum_i p(A_i).
 \end{equation}
+This last requirement codifies the idea that disjoint event are mutually exclusive: if $A$ and $B$ cannot happen at the same time, then the probability of one of the two happening is the sum of them separately happening.
 
-\textbf{Expectation and integration.} TODO: show the use of integrals over probability measure.
+\textbf{Expectation and integration.} A random variable is a function $F : X \to \mathbb{R}$ that assigns a value to each state.\footnote{The function $f$ must be measurable, in the sense that the inverse of a Borel should be a Borel set.
+We find that the term is confusing for a physics audience, so it is avoided in the main text.} The number of the side of the die, position, momentum are typical examples.
+The expectation for a random variable is given by the integral
+\begin{equation}
+	E[F] = \int_X F dp.
+\end{equation}
+The expectation is linear with respect to the variable (i.e.
+$E[aF+bG]= aE[F]+bE[G]$).
+In our case, what is more interesting is that it is a convex function in terms of probability measures: if $p = \lambda p_1 + \bar{\lambda} p_2$ then $\int_X F dp = \lambda \int_X F dp_1 + \bar{\lambda} \int_X F dp_2$.
+That is, as we required for statistical variables, the expectation for a mixture is the mixture of the expectations.
 
-\textbf{Absolute continuity of probability measures.} While each classical ensemble is a probability distribution, not all probability distribution can be thought as a valid ensemble.
-For example, we could have a probability measure to be wholly defined over a single point of phase space, which would correspond to the ability to prepare a system in the same exact place with the same exact momentum.
-Not only this is not feasible experimentally, which violates the whole premise of having objects that map to physically meaningful entities, but it would have to have entropy equal minus infinity.
-To address this, we add a requirement that is both mathematically natural and physically meaningful.
-Let $\mu : \Sigma_X \to [0, + \infty]$ the measure on $X$ that quantifies the count of states in each region.
+\textbf{Absolute continuity of probability measures.} As we saw before, probability measures over a single point over the continuum cannot be understood as valid ensembles.
+To address this, we add a requirement that is both physically meaningful and mathematically useful.
+Let $\mu : \Sigma_X \to [0, + \infty]$ be a measure on $X$ that quantifies the count of states in each region.
 In the discrete case this is simply the counting measure (i.e.
@@
 volumes in phase space correspond to the count of states).
-This measure is additive over mutually exclusive events.
+This measure is additive over disjoint region as they represent mutually exclusive events.
 The requirement is that we can assign non-zero probability only to events that have a non-zero count of states.
@@
 Physically, it means that those points are ``special'' both in terms of experimental verifiability and entropy/count of states.
-While the general framework allows for it, it is not crucial for our discussion here.} This is another example in which a physical requirement (probability must be zero if there are no possible states), once properly spelled out (absolute continuity), justifies common physical intuition (we assume there is a probability density).
+While the general framework allows for it, it is not crucial for our discussion here.} This is another example in which a physical requirement (probability must be zero if there are no possible states), once properly spelled out (absolute continuity), justifies common physical intuition (we assume a probability density exists).
 
-% Intro paragraph that says quickly how to connect to Choquet theory and its limitations. It should be a quick summary of the whole section
-
-\textbf{Choquet theory.} Having reviewed the use of probability measures in classical mechanics, we can ask the more general question: when can we represent ensembles in a generic ensemble space as a probability measure?
-A very similar question is the subject of Choquet theory.
+\textbf{Choquet theory.} Classical statistical mechanics and probability starts by assuming the existence of a state space $X$ and defining ensembles on top of it.
+This cannot be generalized to quantum mechanics and beyond.
+In our theory of ensemble spaces we start with ensembles, and have to show when and how they can be thought, can be represented, as probability measures.
+If we can assume that $E$ is a compact convex set, Choquet theory answers this question.
 
 First, we have to define what it means to represent an ensemble with a probability measure.
-As a first step, let $E$ be a convex subset of a locally convex space $V$.
+Let $E$ be a convex subset of a locally convex space $V$ and $X$ the set of extreme point.
+In classical mechanics, $X$ will correspond to the set of Dirac measures, as these cannot seen as a mixture of other measures.
+In quantum mechanics, $X$ corresponds to the pure states, as these cannot be seen as mixture of other states.
+In other words, the set of states $X$ is recovered as the set of ensembles with the narrowest spread.
 We say that $\ens \in E$ is represented by a probability measure $p : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_E f dp$ for every continuous linear functional $f$ on $V$.
-Physically, instead of describing our ensemble $\ens$ as a point in our ensemble space identified by the value of all possible statistical variables $f(\ens)$, we represent it as a probability measure $p$ and all the associated expectation values.
+Moreover, $p$ is supported by the extreme points if $p(X) = 1$.
+Physically, we start with an ensemble $\ens$ taken as a point in our ensemble space, identified by the value of all possible statistical variables $f(\ens)$, and we represent it as a probability measure $p$, a statistical mixture, over other the extreme points, the states, recovering the expectation with an integral.
 
-In this first step, we are using a probability measure on $E$, so we are defining a probability distribution over ensembles.
-As a second step, we want the probability measure to be defined over the possible cases, the sample space.
-To do that, Choquet theory identifies each possible case with the Dirac measure associated to that case, which is an extreme point of the convex set $E$.
-The result is the following:
-\begin{thrm}[Choquet]
+We should note that, given any probability measure $p$ over the extreme points $X$, we can always find an $\ens \in E$ that is represented by $p$.
+Therefore, we have a map from the set of probability measures to the set of ensembles.
+The first question, then, is whether we can go back.
+Choquet theory tells us that, in the compact case, we can:
+\begin{thrm}[Choquet 1]
 	Let $E$ be a metrizable compact convex subset of a locally convex space $V$, and let $\ens \in E$.
-Then there exists a probability measure $\mu_{\ens}$ on $X$ which represents $\ens$ and is supported by the extreme points of $X$.
+Then there exists a probability measure $p_{\ens}$ on $X$ which represents $\ens$ and is supported by the extreme points $X$.
 \end{thrm}
+Note that an ensemble space is a metrizable convex subset of a locally convex TVS. Therefore, if $E$ is compact, the Choquet theorem applies.
 
-Note that while an ensemble space $\Ens$ is a metrizable convex subset of a locally convex TVS $V$, it is not necessarily compact.
-As we saw, the Dirac measures should not be considered valid ensembles in the continuum case.
-Therefore Choquet theory cannot be applied to our general case.
-When we are drawing the parallel, we will implicitly assume that $E$ is a compact set that is also the closure of $\Ens$ in an appropriate convex TVS $V$.
-Many interesting cases fall into this restriction: the ensemble spaces of classical probability distribution over finitely many discrete cases fall in this restriction; the subset of ensembles over finite ranges of position and momentum; the ensemble spaces of all finite dimensional quantum systems.
-When Choquet theory applies, then, we are able to always characterize an ensemble in terms of a probability distributions over its, maybe idealized, most refined ensembles.
+The above theorem tells us that the map between probability measures and ensembles is surjective. The second question, then, when is it bijective? Again, Choquet theory has a clear answer.
+\begin{thrm}[Choquet 2]
+	Let $E$ be a closed convex metrizable subset of a locally convex space $V$.
+Then $E$ is a Choquet simplex if and only if for each $\ens \in E$ there exists a unique measure $p_{\ens}$ which represents $\ens$ and is supported by the extreme points $X$.
+\end{thrm}
+The Choquet simplex is a generalization of the finite dimensional simplex, whose mathematical characterization is rather technical.
+The key point is that $E$ is a Choquet simplex if and only if the ensemble space does not allow multiple decomposition in terms of pure states.
+This confirms the physical intuition from classical and quantum statistical mechanics.
 
-The previous theorem does not, however, guarantee uniqueness.
-The above theorem, in fact, applies in the quantum case.
-The Bloch ball, in fact, is a compact convex subset and any point can be seen as a convex combination of points on the surface.
-The center, which represent the maximally mixed state, can be thought as an equal mixture of any two opposite points, or a uniform distribution over the whole surface.
-The fact that any ensemble can be seen as a mixture of a set of pure states, then, is not a property of classical spaces.
-The property of classical spaces is that this representation is unique.
-
-Mathematically, the probability measure is unique if and only if $E$ is a Choquet simplex, which is a generalization of the finite dimensional simplex.
-The characterization of the Choquet simplex is fairly technical, so we will simply define the Choquet simplex by the property of having a single decomposition in terms of extreme points.
-The classical case, that is when the ensemble space coincide with the space of all probability measures over a set of states, is a Choquet simplex.
-
-% space is metrizable https://proofwiki.org/wiki/Birkhoff-Kakutani_Theorem/Topological_Vector_Space
+For example, the ensemble space for a two state quantum system, the Bloch ball, is a compact convex subset.
+Any internal point can be seen as the convex combination of points on the surface, the pure states.
+The center, which represents the maximally mixed state, can be thought as an equal mixture of any two opposite points, or a uniform distribution over the whole surface.
+The fact that any ensemble can be seen as a mixture of a set of pure states, then, is reflected by the first part of the Choquet theorem, and it is valid for all physical theories.
+Uniqueness of the representation fails, however, precisely because the ensemble space is not a simplex: it is not a classical space.
 

### 2025-07-21 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed quasi-probability

@@
 \begin{equation}
-	W(\rho) = W_{\hat{\rho}} (q, p) = 2 \int_{-\infty}^{\infty} \langle x + y | \rho | x - y \rangle e^{-2 i p y / \hbar} \; \mathrm dy,
+	W(\rho) = W_{\rho} (q, p) = 2 \int_{-\infty}^{\infty} \langle x + y | \rho | x - y \rangle e^{-2 i p y / \hbar} \; \mathrm dy,
 \end{equation}
-takes a mixed state $\rho$ and returns a function of phase space that integrates to one, like a probability density, but can have small negative regions.
+takes a mixed state $\rho$ and returns a real function $W_{\rho}$ of phase space that, like a probability density integrates to one, but, unlike a probability density, can be negative.
 Recall that an affine combination is a linear combination where all coefficients sum to one but are not necessarily non-negative.
 Therefore, in general, quasi-probability distributions can be understood as affine combinations.
-Remarkably, expectation values of polynomial of position and momentum in symmetric ordering, coincide with expectations over the Wigner function.
+Remarkably, expectation values of polynomial of position and momentum in symmetric ordering coincide with expectations over the Wigner function.
 For example
@@
 Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events, and all mutually exclusive events are represented by disjoint regions.
-Mutually exclusive events and be prepared and observed independently from the others (i.e.
+Mutually exclusive events can be prepared and observed independently from the others (i.e.
 they are orthogonal in the sense that they maximize entropy increase during mixture); we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasi-probability in quantum mechanics.
@@
 where $F_\psi$ is an entire analytic function, which cannot vanish on an open set without vanishing everywhere, therefore the support of all Husimi Q functions is the whole space $Z$.
-This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support; moreover disjoint subsets of $Z$ cannot be thought of as mutually exclusive events because event limited to any proper subset exists.
+This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support; moreover disjoint subsets of $Z$ cannot be thought of as mutually exclusive events because we are not going to be able to find two distribution with those subsets as their respective supports.
 For the discrete case, a SIC-POVM representation picks $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot be all orthogonal to each other.
@@
 Therefore, disjoint sets of $P_i$ do not correspond to mutually exclusive events.
-Pseudo-probability representation, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
-
-While negative probability is a red herring, we want to understand two things.
-When can we use classical probability?
-Are quasi-probability representation specific to quantum mechanics, or are they more general?
-Fortunately, the ensemble space framework allows us to answer these questions directly.
+Quasi-probability representation, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
 
@@
 
-\textbf{Ensemble space embedding and signed measures.} As we saw when reviewing Choquet theory, and measure on the extreme points represents an ensemble.
-Classical spaces are those in which there is a one-to-one correspondence between each measure and each ensemble.
-We also saw that the hallmark of pseudo-probabilities is the lack of mutual exclusivity of the events.
-Why is mutual exclusivity linked to the lack of multiple representations?
+\textbf{Ensemble space embedding and signed measures.} We saw that negative probability is a red herring, and mutual exclusivity is key.
+But how exactly mutual exclusivity give us classical probability?
+Moreover, are quasi-probability representations specific to quantum mechanics, or are they more general?
+This is where the perspective from the ensemble space framework allows us to answer these questions directly.
 
+In Choquet theory, we saw that classical spaces, Choquet simplexes, are those in which there is a one-to-one correspondence between probability measures and ensembles, where no multiple decomposition in terms of pure state is allowed.
+We want to show that this is a consequence of assuming mutual exclusivity between the points of $X$.
 Suppose we have an ensemble $\ens$ that can be represented as different measures $p_1$ and $p_2$ over the extreme points.
-This means that there are two sets $A$ and $B$ for which $p_1(A) < p_2(A)$ and $p_1(B) > p_2(B)$.
+This means that there are two disjoint sets $A, B \subseteq X$ for which $p_1(A) < p_2(A)$ and $p_1(B) > p_2(B)$.
 That is, we can move probability from $A$ to $B$ while leaving the ensemble, and therefore the entropy and all statistical variables, unchanged.
@@
 The only way for $A$ and $B$ to be mutually exclusive, then, is if different measures represent different ensembles, which means the ensemble space is a simplex.
-This answers the first question: we can use classical probability exactly when all extreme points are mutually exclusive (i.e.
+This tells us that we can use classical probability exactly when all extreme points are mutually exclusive (i.e.
 orthogonal as defined from the entropy) and the ensemble space is a simplex.
 
+That is, classical probability does not merely correspond to the use of a single non-negative probability measure.
+It corresponds to the ability to, conceptually, allow any possible probability measure to describe a possible distinct scenario.
+This requirement is exactly mutual exclusivity.
 In quantum mechanics, a classical context is a simplex embedded within the ensemble space.
-For the Bloch ball, a classical context is an axis that connect two opposite point.
+For the Bloch ball, a classical context is an axis that connects two opposite point.
 The points are orthogonal because the midpoint is the maximally mixed state, which therefore maximizes the entropy increase, and the line segment is a one simplex.
-If we take the position operator and consider all density operators that commute with $x$, this is exactly functions $\rho(x)$ that integrate two once, which is the space of probability supported by $X$, which is a simplex.
-We want to stress how the conceptual view and the mathematical view fits precisely together with little effort.
+If we take the position operator and consider all density operators that commute with $x$, this is corresponds to the set of functions $\rho(x)$ that integrate two once, which is the space of probability supported by $X$, which is a simplex.
 
-For the second question, let $Z$ be the space over which we are defining our quasi-probability distributions.
+Now the second question: why does quantum mechanics allow quasi-probability representations?
+Let $Z$ be the space over which we are defining our quasi-probability distributions.
 Note that, in general, $Z$ is not the space of states.
@@
 
-\textbf{Non-uniquess of quasi-probability.}
-
-This alternative way of looking at measure theoretic representations, then, is consistent with the previous one and fits in our general framework.
-Since quantum ensemble spaces are not simplexes, classical probability does not apply.
-The different bounds one gets between classical and quantum mechanics depend on the different shape of the respective ensemble spaces.
+\textbf{Non-uniquess of quasi-probability.} Our approach gives us a consistent view of both quasi-probabilities and classical probability.
+The first are affine representations of the ensembles that connect to the vector space structure, while the second are convex representation of the extreme points which rely on the shape of the ensemble space.
 
 Quasi-probability distributions are very useful in practice, particularly because one can construct them to fit particular problems.
-Once one has identified a set of observables that are the most natural ones to define a problem, one can find a representation that is specifically suited for the case.
-While the quasi-probability cannot be interpreted as probability, it still provides a useful tool to describe and understand the problem.
+Once one has identified a set of observables that are the most natural ones to define a problem, one may find a representation that is specifically suited for the case.
+While the quasi-probability cannot be interpreted as probability, it still provides a useful tool to describe and understand the problem and give some intuition.
 
 However, this practical convenience is also a drawback for finding a more general approach to probability in spaces beyond classical mechanics.
-Crucially, the space $Z$ over which the quasi-probability measure are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tells us which quasi-probability measure represent valid ensembles and which do not.
+Crucially, the space $Z$ over which the quasi-probability measures are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tells us which quasi-probability measure represent valid ensembles and which do not.
 
 As a comparison, in the previous case all probability measures were given over a uniquely defined set of points, the extreme points of the ensemble space, but the probability measure may not be unique.
-In this case, the pseudo-probability associated to each ensemble is unique, but the space over which the pseudo-probability is defined is not.
+In this case, the quasi-probability associated to each ensemble is unique, but the space over which the quasi-probability is defined is not.
 

### 2025-07-21 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed capacities

@@
 
-\textbf{Probability of outcomes vs fraction of components.} From the previous discussion, we can see that the problem with using probability in quantum mechanics is that not all pure states represent mutually exclusive events, and it is only over mutually exclusive events that classical probability can be defined.
-Mutually exclusive events correspond to ensemble are orthogonal, as each instance of one ensemble cannot be produced by the other ensemble.
+\textbf{Probability of outcomes vs fraction of components.} We saw that classical probability fails in quantum mechanics because pure states are not always mutually exclusive, are not orthogonal.
 That is why, in quantum mechanics, probability can be defined only once a context, an orthogonal basis, is chosen.
 Therefore, at a conceptual level, probabilities can only give us a patchwork representation of ensembles.
-Some author embrace this view, by characterizing states as the collection of all probability distributions of all possible results, as it is generally the case in GPTs.
+Some author embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.
 [CITATION NEEDED] However, there is a different route.
@@
 That is, regardless of whether we can characterize experiment outcomes with probability distributions, preparation procedures will always allow us to create statistical mixture.
-We can create statistical mixtures regardless of whether the states we are mixing represent mutually exclusive outcomes or not.
-That is, instead of focusing on the likelihood of an outcome, we will focus on how much of a particular ensemble can be understood as the mixture of some other ensembles.
-Instead of focusing on what an ensemble may cause, we focus on how an ensemble can be produced.
-It turns out, this second approach generalizes in a natural way.
+As we saw in Choquet theory, the difficulty is that there are multiple was to mix pure states to recover the same ensemble.
+However, the question of what is the biggest mixture coefficient one can use in a specific case is well-defined, and generalizes to the non-compact case as well.
+Therefore, our generalization of probability is not in terms of what an ensemble can lead to (i.e.
+measurements), but in terms of how ensembles can be produced (i.e.
+preparations).
 
@@
 The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
-fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\footnote{ \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} } We will see how our approach fits  in that literature in a later section.
+fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\footnote{ \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} }
 
-\textbf{Compare to Choquet theory.} There is a straight forward link between our non-additive measure and Choquet theory.
-Recall that Choquet theory start by defining measure on the whole ensemble space before finding those constrained to the extreme points.
-Our setup is similar.
-The fraction capacity is defined on the whole ensemble space.
-In the classical discrete case, for example, if $A \subset \Omega \subset \Ens$ is an event, then $\frcap_{\ens}(A)$ will return the probability of event $A$; if $P \subset \Ens$ represents all the Poisson distributions with integer coefficient, $\frcap_{\ens}(P)$ will return how much of $\ens$ can be expressed as a mixture of Poisson distributions.
-In case that extreme points exist, or we can extend the ensemble space with some limit procedure, we can also define the fraction capacity on the extreme points.
-
-We can establish a direct connection between the fraction capacity and Choquet theory, because the ability to express an ensemble $\ens$ as the convex combination of $A$ is precisely the ability to find a measure $\mu$ that has support within $A$.
-That is, $\ens = \sum_{i=1}^{\infty} p_i \ens[a]_i$ if an only if the probability measure defined by $\mu(\{\ens[a]_i\}) = p_i$ represents $\ens$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{\mu(A) \, | \, \mu \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all possible measures.
+\textbf{Compare to Choquet theory.} In case $E$ is compact, we can restrict the fraction capacity to the set $X$ of extreme points.
+Given a set $A \subset X$, of extreme points, we can find the set $N \subset E$ of all ensemble that can be represented with a probability measure with support $A$.
+This forms a convex set.
+Now, the ability to express an ensemble $\ens$ as the convex combination of $N$ is precisely the ability to find a measure $p$ that represents $\ens$ such that $p(A) \neq 0$.
+That is, $\ens = \lambda \ens[a] + \bar{\lambda} \ens[b]$ with $\ens[a] \in M$ if and only if there is a probability measure $p$ that represents $\ens$ such that $p(A) \geq \lambda$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{p(A) \, | \, p \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all possible measures.
 
@@
 In the non-classical case, the closure of the ensemble space will not be a simplex, and there will be multiple representations, which will lead to a non-additive measure.
-Suppose, in fact, that $\mu$ and $\nu$ are two distinct measures that represent the same ensemble $\ens$.
-Then there will be a set $A$ such such that $\mu(A) \neq \nu(A)$.
-Suppose, without loss of generality, that $\mu(A) > \nu(A)$ We will also have $\mu(A^{\complement}) = 1- \mu(A) < 1 - \nu(A) = \nu(A^{\complement})$.
-Since the fraction capacity for a set cannot be less than the measure on that set, we have $\frcap_{\ens}(A \cup A^{\complement})=\frcap_{\ens}(X) = 1 = \mu(A) + \mu(A^{\complement}) < \mu(A) + \nu(A^{\complement}) \leq \frcap_{\ens}(A) + \frcap_{\ens}(A^{\complement})$.
+Suppose, in fact, that $p_1$ and $p_2$ are two distinct measures that represent the same ensemble $\ens$.
+Then there will be a set $A$ such such that $p_1(A) \neq p_2(A)$.
+Suppose, without loss of generality, that $p_1(A) > p_2(A)$ We will also have $p_1(A^{\complement}) = 1- p_1(A) < 1 - p_2(A) = p_2(A^{\complement})$.
+Since the fraction capacity for a set cannot be less than the measure on that set, we have $\frcap_{\ens}(A \cup A^{\complement})=\frcap_{\ens}(X) = 1 = p_1(A) + p_1(A^{\complement}) < (A) + p_2(A^{\complement}) \leq \frcap_{\ens}(A) + \frcap_{\ens}(A^{\complement})$.
 That is, the fraction capacity is additive over disjoint sets of extreme points in and only in the classical case.
@@
 \textbf{Uniqueness of the fraction capacity.} From a conceptual standpoint, the fraction capacity has a major advantage: it has a clear physical meaning and it is unique.
-Like in Choquet theory, and unlike for pseudo-probability, the domain over which the measure is defined is fully determined by the system, by the ensemble space.
-However, unlike Choquet, and like pseudo-probability once the space is fixed, there one unique fraction capacity for each ensemble.
+Like in Choquet theory and unlike for quasi-probability, the domain over which the measure is defined is fully determined by the system, by the ensemble space.
+On the other hand, unlike Choquet theory and like quasi-probability, once the space is fixed, there is one unique fraction capacity for each ensemble.
 Moreover, the setup is more general than Choquet theory (i.e.
@@
 To build an intuition, consider the classical case.
-For every a region $U$ of phase space, we have a set of possible probability distributions that have support $U$.
+For every region $U$ of phase space, we have a set of possible probability distributions that have support $U$.
 This is a convex set since every mixture of distributions with support $U$ has support $U$.
@@
 If $A$ happens to be the set of ensembles with support $U$, then the supremum of the entropy is $\log \mu(U)$, which works also in the unbounded case.
-But the setup now is more general: $A$ can be any set of ensembles in any ensemble space.
 
-We still need to make sure that this definition still makes sense physically.
-If the entropic volume measure the spread of an ensemble, during a mixture the spread can at most sum.
+The above intuition started with the notion of the state space $X$ and the count of states $\mu$.
+In the general case, we will start with the ensemble space $\Ens$ and the entropy $S$.
+For ensembles, we can define the entropic volume\footnote{CITE Hall} to be the exponential of the entropy $2^{S(\ens)}$.
+For the entropic volume to quantify the spread of the ensemble, we must have that the spread can at most sum.
 Remarkably, the bounds on the entropy are such that the exponential of the entropy is subadditive.
-That is, $2^{S(p\ens[a]+\bar{p}\ens[b])} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])}$ for any ensemble space.\footnote{It is an open question whether the exponential is the only monotonic function of the entropy that is subadditive.} Therefore this is indeed a general way to define the spread of ensembles and therefore the count of distinguishable states.
+That is, $2^{S(p\ens[a]+\bar{p}\ens[b])} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])}$ for any ensemble space.\footnote{It is an open question whether the exponential is the only monotonic function of the entropy that is subadditive.} For the equality to hold, the ensemble must be orthogonal, meaning that two states are distinguishable if and only if they are mutually exclusive.
+Therefore, this is indeed a general way to define the spread of ensembles and the count of distinguishable states in a way that is consistent with the previous discussions.
 
@@
 Again we note that the non-additive measure we recover is very well-behaved, and connects directly with important quantities.
-When Choquet theory applies, state capacity can be extended to the extreme points.
+In the compact case, state capacity can be extended to the extreme points.
 In the classical case, disjoint sets of extreme points will be orthogonal, meaning that the state capacity there restricted is an additive measure.
@@
 In quantum mechanics, only orthogonal state are distinguishable, therefore the measure is non-additive over pure states.
-We see again how the lack of single decomposition of ensembles over pure states is responsible for the failure of the additivity of the measure.
+We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutually exclusiveness, is responsible for the failure of the additivity of the measure.
 

### 2025-07-22 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed last section

@@
 
-\textbf{Requirements.} In a previous section we reviewed what types of objects we have in probability theory for classical ensembles: a sample space $X$, a probability measure $p$, random variables $F : X \to \mathbb{R}$, expectations as integrals of the probability measure $E[F] = \int_X F(x) dp$, a state counting measure $\mu$, the probability density as the Radon-Nikodym derivative $\frac{dp}{d\mu}$ and the entropy as $\int_X \rho \log \rho d\mu$.
+\textbf{Requirements for non-additive probability theory.} In a previous section we reviewed what types of objects we have in probability theory for classical ensembles: a sample space $X$, a probability measure $p$, random variables $F : X \to \mathbb{R}$, expectations as integrals of the probability measure $E[F] = \int_X F(x) dp$, a state counting measure $\mu$, the probability density as the Radon-Nikodym derivative $\frac{dp}{d\mu}$ and the entropy as $\int_X \rho \log \rho d\mu$.
 Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the states counting measure $\mu$, is there a generalization of all other elements?
@@
 The first difficulty, as we already pointed out, is that the ensemble space $\Ens$ may not have extreme points.
-What is the correct procedure to ``close'' the ensembles space and recover the space of extreme points is, for us, an open question.
-\footnote{One attempt, for example, was by using orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
+What is the correct procedure to ``close'' the ensembles space and recover the space of extreme points is, for us, an open question.\footnote{One attempt, for example, was by using orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
 However, this fails to recover the points correctly on the continuum even for classical mechanics.
-We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} For the purpose of this paper, we will assume $\Ens$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
-Therefore we can talk about the restriction on $X$ of the state capacity, the fraction capacity and the statistical variables.
+We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} For now, we will assume $E$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
+Therefore we can talk about the restriction on $X$ of the state capacity and the fraction capacity.
 
-Recall that a statistical variable is linear function $F : \Ens \to \mathbb{R}$ defined over the ensemble space.
-Given that $F$ is a linear, its restriction on $X$ will determine the values for all other points of $\Ens$.
+Recall that a statistical variable is linear function $F : E \to \mathbb{R}$ defined over the ensemble space.
+Given that $F$ is a linear, its restriction on $X$ will determine the values for all other points of $E$.
 Since the expectation is recovered through the integration of the variable over the probability density, we would like a non-additive integral $(E)\int_X$ such that
@@
 Depending on the use cases, different types of integrals exists.
-In fact, procedures exist\cite{nonadditive2014,pap2013} to construct infinitely many types of non-additive integrals.\footnote{Briefly, one can define a pseudo-addition and a pseudo-multiplication as real functions of two real arguments and construct an integral based on those definitions.} We can, however, at least rule out the two most common non-additive integrals: the Sugeno integral and the Choquet integral.
+In fact, procedures exist\cite{nonadditive2014,pap2013} to construct infinitely many types of non-additive integrals.\footnote{Briefly, one can define a pseudo-addition and a pseudo-multiplication as real functions of two real arguments and construct an integral based on those definitions.} At this point, we can rule out the two most common non-additive integrals: the Sugeno integral and the Choquet integral.
 
@@
 
-The ensemble space $\Ens$ for a two-state quantum system is the Bloch ball.
+The ensemble space $E$ for a two-state quantum system is the Bloch ball.
 The pure states, the extreme points, are on the surface $X$ and the internal points correspond to mixed states.
@@
 the eigenstates) will take the minimum and maximum values $a = F(\ens[a])$ and $b = F(\ens[b])$.
-For simplicity, we can picture $\ens[a]$ and $\ens[b]$ as the extreme points of the vertical axis.
+For simplicity, we can picture $\ens[a]$ and $\ens[b]$ as the extreme points of the vertical axis and we will assume $0<a<b$.
 
@@
 Also note that the distance between $\ens[o]$ is the same for all extreme points, therefore, to maximize $p$, we simply need to minimize the distance between $\ens[o]$ and $\ens$.
-We have that $\ens$ must be the intersection between the vertical axis and the disk, and $\ens[c]$ will coincide with $\ens[a]$.
-The coefficient $p$ will be given by ratio of the distance $\overline{\ens[o]\ens[a]}$ and $\overline{\ens\ens[a]}$.
+Therefore $\ens$ must be the intersection between the vertical axis and the disk, and $\ens[c]$ will coincide with $\ens[a]$.
+The coefficient $p$ will be given by the ratio of the distance $\overline{\ens[o]\ens[a]}$ and $\overline{\ens\ens[a]}$.
 Since $F$ is a linearly functional, we have
@@
 As we saw before, the additivity of the measure corresponds to the ability to assign probability to a set without implicitly assigning probability to another.
-In terms of statistical variable, two mutually exclusive events can be assigned values independently.
+In terms of statistical variables, two mutually exclusive events can be assigned values independently.
 This, again, recovers how observable work in quantum mechanics, without mentioning quantum mechanics: only orthogonal states, that is mutually exclusive events, can associated to measurement outcomes and be given a classical probability distribution.
 The non-additivity of the fraction capacity and the non-arbitrariness of the statistical variables over the extreme points, then, must be linked to the shape of the ensemble space.
-To our knowledge, there is no such requirement for the more typical used of non-additive measure theory.
 
@@
 Therefore state capacity, fraction capacity and statistical variables have to be, in a sense, compatible with each other.
-
-As we saw in the classical case, probability measures have to be absolutely continuous with respect to the state counting measure.
-Also, if the state capacity is additive, all disjoint events are mutually exclusive, which means the fraction capacity is additive and disjoint events can correspond to independent values for a statistical variable.
-Therefore, we wonder whether a more general compatibility condition exists between the fraction capacity and the state capacity/statistical variable respectively.
-
-We have not found similar discussion in the literature for non-additive measure, which makes us suspect that these considerations are specific to this setting.
+One may be able to use the state capacity to characterize the degree of additivity of the fraction capacity.
+Since in non-additive measure theory, family of capacities are often characterized the type of non-additivity,\footnote{For example, a Sugeno $\lambda$-measure is a non-additive measure $\mu$ such that $\mu(A \cup B) = \mu(A) + \mu(B) + \lambda \mu(A)\mu(B)$ for some $\lambda$ over disjoint sets $A$ and $B$.
+If $\lambda = 0$, the measure is additive.} this may give a way forward.
+Moreover, we saw that in the classical case probability measures have to be absolutely continuous with respect to the state counting measure.
+Maybe an elegant and meaningful compatibility condition exists between the fraction capacity and the state capacity/statistical variable that captures all requirements.
+Nothing along these lines came up in our literature search.
 

### 2025-07-28 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: minor changes

@@
 
-\textbf{Counting states.} We now want to introduce a measure theoretic representation of the entropy that generalizes the notion of count of states.
+\textbf{Counting states.} Since the fraction capacity can be seen as the non-additive generalization of a probability measure, we now want to introduce a similar generalization for the notion of count of states.
 Since entropy characterizes the variability of the preparations of the ensemble, greater variability means the ensemble is spread over more cases.
@@
 The state capacity, then, recovers the number of distinguishable states.
-In classical mechanics, all states are perfectly distinguishable, the measure is additive.
+In classical mechanics, all states are perfectly distinguishable as they represent mutually exclusive cases, the measure is additive.
 In quantum mechanics, only orthogonal state are distinguishable, therefore the measure is non-additive over pure states.
 We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutually exclusiveness, is responsible for the failure of the additivity of the measure.
+Note that, since the fraction capacity depends only on the ensemble space, and not also on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
 
@@
 The first difficulty, as we already pointed out, is that the ensemble space $\Ens$ may not have extreme points.
-What is the correct procedure to ``close'' the ensembles space and recover the space of extreme points is, for us, an open question.\footnote{One attempt, for example, was by using orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
-However, this fails to recover the points correctly on the continuum even for classical mechanics.
-We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} For now, we will assume $E$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
+What is the correct procedure to ``close'' the ensembles space and recover the space of extreme points is, for us, an open question.
+For now, we will assume $E$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
 Therefore we can talk about the restriction on $X$ of the state capacity and the fraction capacity.

### 2025-07-28 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Replace sketch
Replaces the hand-drawn sketch for the Choquet integral calculation with a TikZ picture.
NonAdditiveProbability: reviewed the introduction

@@
 
-\section{TOC}
-
-Overall goal: to create a generalization of probability theory and measure theory for general physical systems.
-
-General question: how do we represent physical states? Different types of representations: talk about examples, ....
-
--- General problem
-
-For this work: assume we have a locally convex topological vector space V. E is a bounded convex subset of V. Equipped with an entropy. Topology generated by entropic balls?
-
-Separateness and orthogonality
-
-Statistical quantities
-
-Show that they reduce to standard things in classical mechanics and quantum mechanics.
-
--- Connection to Choquet theory
-
-Choquet theory - in the case that the closure of the space is compact -> representation through measures. May not be unique
-
-Choquet simplex. Uniqueness of decomposition and separateness/orthogonality.
-
--- Connection to signed probability
-
-TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
-
-Characterize signed measures as an affine representation of the ensemble space (i.e. bounded convex subset of a locally convex topological vector space).
-
-Show that the probability is signed if and only if multiple decomposition, non-additive, no Choquet simplex
-
-Signed probability is unique, but does not characterize the convex set (ensemble space).
-Note: the other measures are defined on the extreme points.
-Knowing the extreme points means you know the convex space (i.e.
-take the hull).
-The non-additive measure does not characterize the extreme points, so it doesn't really characterize the space.
-
--- Our non-additive measure approach
-
-State capacity
-
-Fraction capacity. Show uniqueness. Also show uniqueness on extreme points.
-
-Fraction capacity as supremum of all possible measures
-
-Additivity of fraction capacity (and state capacity).
-
-Single decomposition and classical spaces and measurement contexts
-
-Show that they reduce to standard things in classical mechanics and quantum mechanics.
-
--- Connection to fuzzy logic?
-
-Literature search.
-
-Possible connection to
-* Choquet integral
-* Sugeno integral
-
-? Space of fuzzy measure like we have a space of additive measures?
+%TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
 
@@
 
-In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories.
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts, which stems from the underlying aim to characterize physical theories from an operational viewpoint.
+
+in terms of operations aspects, which lends themselves ,  Given the statistical/probabilistic nature of measurement outcomes and the success of information theory, a generalized notion of probability is typically assumed.
+.
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
@@
 
-We reformulate the problem in the following way: given the space of all possible statistical ensembles allowed by a statistical theory, what measure theoretic tools can we use to represent each ensemble and the geometric properties of the overall space?
-Why exactly classical probability measures fail to capture completely quantum states?
-What suitable generalizations are possible, and what mathematical tools would need to be developed?
+The paper aims to ask the following question: can we, by arguing from necessity, find a minimal set of physical requirements for a state space of a physical system, that allows to directly connect to all the measure theoretic structures currently used in physics (i.e.
+probability measures, Liouville measure to count states, quasi-probability, ...) and point to a straight-forward generalization of such structures?
+We will show that this is indeed possible.\footnote{Note that we believe that the requirements we set, though necessary, are not sufficient to generalize the Lie algebra given by the classical Poisson brackets and the quantum commutators, meaning they cannot fully recover mechanics.
+Nonetheless, they are enough to recover measure theoretic structures, the topic of this paper.}
 
-In this paper, we will characterize a generic space of ensembles as a convex bounded set of a Hausdorff locally convex topological vector space, upon which an entropy function is defined.
-These mathematical properties are physically grounded in the ability to experimentally define ensemble (leading to a topological space), to define statistical mixtures (leading to a convex space), to quantify the variability of the elements of the ensembles (leading to an entropy function) and to having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
+The starting point is that any physical law deals with experimentally reproducible relationships.
+This implies a physical theory must gives us the space of statistical ensembles (i.e.
+the collection of all possible preparation procedures) that the theory allows.
+As a basic requirements, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from GPTs is the entropy requirement.
+We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
+Mathematically, under these premises, an ensembles space is a convex subset of a locally convex topological vector space.
 The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the dimension of orthogonal subspaces of an operator)?
@@
 
-We will first review some results from Choquet theory, which is, to our knowledge, the most general approach to represent elements of a convex set using additive probability measures.
-The key insight is that, at least in the compact case, such representation is always possible, but it is not unique.
-Uniqueness requires the space to be a Choquet simplex, which is a generalization of the standard finite-dimensional simplex and is a particularly shaped subset of a vector space.
-The space of quantum ensembles is not a simplex for one key reason: it allows multiple decompositions in terms of pure states.
-These insights will allow to better understand exactly the limits of classical probability to describe ensembles in physical theories.
+We will first review how measure theory is used in classical probability together with some results from Choquet theory.
+This will tell us that elements of a compact convex set can always be represented by a probability measure over its extreme points, the pure states.
+While such representation is always possible, at least in the compact case, it is not unique.
+Uniqueness requires the space to be a Choquet simplex, which is a generalization of the standard finite-dimensional simplex and represents the classical case.
+The key insight, then, an ensemble space is classical if and only if each ensembles admits a unique decomposition in terms of pure states.
 
-We will also review some results from the use of quasi-probability measures.
-While the Wigner function represents the most known instance, there are different choices.
-Note that the space of signed probability measures over a set $K$ is a vector space, so one can find an affine map between a suitable space of signed probability measures and the embedding vector space.
-Each ensemble, then, is uniquely represented with a signed probability measure.
-However, unlike in the Choquet case, the measure is not supported by the set of extreme points of the ensemble space.
-That is, there is no unique or privileged choice for this representation.
-Moreover, since the map is at the level of the full vector space, not just the subset with valid ensembles, there is no general way to characterize which subset of signed probability measures correspond to ensembles.
+We will then review some results from the use of quasi-probability measures.
+While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.
+The key insight is that it is not negative probability that characterizes quasi-probability, but the lack of mutual exclusivity of the points of the space.
+The quasi-probability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
 Therefore, while signed measures are a powerful tool for calculation in specific problems, they cannot give us a general principled representation of the physical theory.
@@
 The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
-From the entropy, we can define the state capacity, which generalizes the notion of count of states.
-For each ensemble, we can define the fraction capacity, which generalizes the notion of probability.
-Both are non-negative sub-additive monotonic set functions, with the second being unit bounded.
-These measures are always unique and, since they are defined on the ensemble space itself, always have provide a physically meaningful representation.
-Moreover, in the compact case, these measure can be restricted on the extreme points while retaining uniqueness.
-(TODO: prove)
+Given an ensemble $\ens$, we can ask what fraction can be produces by a mixture of a given ensembles.
+This leads to the fraction capacity: a sub-additive set function that, like the probability measure, is bounded to one, monotonic and continuous from above and below.
+Over the extreme points, the fraction capacity represents the supremum of all possible measures that can represent the same ensemble in Choquet theory.
+The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure, and is therefore additive.
+This again solidifies the connection between classical probability, additivity of measure and mutual exclusivity.
 
-The fraction capacity is the supremum of all possible measures for the same ensemble.
-The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure, and is therefore additive.
-The state capacity, instead, corresponds to the minimum of the entropy of all possible representations.
-It recovers the Liouville measure in the classical case and the dimensionality of the Hilbert space in the quantum case.
+We can also generalize a notion of count of states.
+While in classical statistical mechanics we start with a measure that quantifies the count of states and define the entropy using its logarithm, for the ensemble space we proceed in the opposite way: we start from the entropy and we define the count of states as the exponential of the entropy (i.e.
+the entropic volume).
+From this, we can define a state capacity, which is again a sub-additive measure.
+It reduces to the Liouville measure in the classical case and the dimensionality of the Hilbert space in the quantum case.
 
-Having shown that a general approach based on non-additive measures is possible, we lay down a series of questions to explore in the future.
-The biggest one is whether it is possible to generalize notions such as integration, derivation and expectation to this non-additive case.
-It should be noted that non-additive measures are used in the context of decision theory, so there may be a connection between the use of non-additive measures in quantum theory and interdependent complimentary choices.[TODO: find citation that sub-additive is when choices are complementary] Both the Choquet and the Sugeno integral, for example, are defined over a monotonic set-functions, so there is a chance that these tools may turn out to be useful.
-We see this work as starting this line of research, which will take more effort to investigate.
+Lastly, we lay down the requirements for a non-additive generalization of probability theory.
+While should be in principle possible, we show that the most used non-additive integral, the Sugeno and Choquet integral, do not work for our purposes.
+This means that more work will be needed to develop the appropriate tools.
 

### 2025-07-28 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: obvious typos

@@
 a positive semi-definite self-adjoint operator of trace one).
-A physical law, then, ultimately expresses relationship among ensembles.
+A physical law, then, ultimately expresses relationships among ensembles.
 A minimal requirement for a physical theory, then, is to provide an ensemble space: the collection of all possible ensembles within the theory, together with a basic mathematical structure that all physical theories must possess.
@@
 
-Additionally, we will require that ensembles can be identified through real valued statistical quantities.\footnote{It is an open question whether this is an additional assumption or it can be derived from the other requirements.} A statistical quantity represents the expectation of a real valued quantity, and is therefore an affine continuous functions $F : \Ens \to \mathbb{R}$.
+Additionally, we will require that ensembles can be identified through real valued statistical quantities.\footnote{It is an open question whether this is an additional assumption or it can be derived from the other requirements.} A statistical quantity represents the expectation of a real valued quantity, and is therefore an affine continuous function $F : \Ens \to \mathbb{R}$.
 That is, $F(p \ens[a] + \bar{p} \ens[b]) = p F(\ens[a]) + \bar{p} F(\ens[b])$.
@@
 
-Each physical theory satisfies these axioms, even though it may be represented very different mathematically.
+Each physical theory satisfies these axioms, even though it may be represented very differently mathematically.
 Classical mechanics, for example, is typically described in terms of the geometry of phase space.
@@
 Therefore, while the above requirements are needed for a physical theory, it does not necessarily mean that it is convenient to describe problems or make calculations in terms of those definitions.
-In fact, those requirements interact with each other leading to a rich mathematical structure that connect to many of the structures we use in mathematical physics, and its goal is mainly to see how the different representations are connected.
+In fact, those requirements interact with each other leading to a rich mathematical structure that connects to many of the structures we use in mathematical physics, and its goal is mainly to see how the different representations are connected.
 
-The goal of this paper is to examine the different way that measure theory can be used to represent ensembles in any given physical theory.
+The goal of this paper is to examine the different ways that measure theory can be used to represent ensembles in any given physical theory.
 
@@
 
-\textbf{Ensemble spaces as convex sets.} For the purpose of measure theoretic representations, it is best to understand an ensemble space $\Ens$ as a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{The entropy constraints forces the topology to be $\mathsf{T}_1$ the convex space to embed into a vector space.
+\textbf{Ensemble spaces as convex sets.} For the purpose of measure theoretic representations, it is best to understand an ensemble space $\Ens$ as a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{The entropy constraint forces the topology to be $\mathsf{T}_1$ the convex space to embed into a vector space.
 A second countable $\mathsf{T}_1$ TVS is metrizable.
-The family of statistical quantities induce a family of semi-norms on the space, which must be countable due to second countability, which makes the TVS locally convex.} The entropy is a function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
+The family of statistical quantities induces a family of semi-norms on the space, which must be countable due to second countability, which makes the TVS locally convex.} The entropy is a function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
 \begin{equation}
@@
 However, on physical grounds, these may not be proper ensembles.
-In classical mechanics, these would corresponds to Dirac measure: probability distributions all concentrated at a single point of phase space.
-Since we cannot reliably prepare a particle with infinite precision, and its entropy would correspond to minus infinity, these are not ensemble and they are not part of $\Ens$.
+In classical mechanics, these would correspond to Dirac measures: probability distributions all concentrated at a single point of phase space.
+Since we cannot reliably prepare a particle with infinite precision, and its entropy would correspond to minus infinity, these are not ensembles and they are not part of $\Ens$.
 For the purpose of Choquet theory, we will assume that there is a way to find the closure $E \supset \Ens$ that includes all extreme points $X$.
 Moreover, we will also assume $E$ to be compact, even though it will not be in general.
-How to properly make these identification, and how to go beyond the compact case, is outside of the scope of this particular article.\footnote{An early attempt used the notion of orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
+How to properly make these identifications, and how to go beyond the compact case, is outside of the scope of this particular article.\footnote{An early attempt used the notion of orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
 This does not work.
@@
 We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} Note that the classical discrete and the quantum case fall in the compact case.
-Moreover, the set of all probability distributions over a compact space $X$ also fall in the compact case.
+Moreover, the set of all probability distributions over a compact space $X$ also falls in the compact case.
 Therefore, apart from the technicality above, the compact case is still a very interesting case.
@@
 This will recover the notion of orthogonality for both classical and quantum spaces.
-Conceptually, these corresponds to two ensembles where all instances are distinct from the other: they are mutually exclusive and a single instance is enough to distinguish between the two cases.
+Conceptually, these correspond to two ensembles where all instances are distinct from the other: they are mutually exclusive and a single instance is enough to distinguish between the two cases.
 
-In classical spaces, orthogonality and separatedness coincide: two ensembles are orthgonal and separate if and only if the support of the respective probability distributions overlap on a non-measure-zero set.
+In classical spaces, orthogonality and separatedness coincide: two ensembles are orthogonal and separate if and only if the support of the respective probability distributions overlap on a non-measure-zero set.
 In quantum mechanics, these two notions are different: any two pure states are separate but are not orthogonal.
@@
 That is, we are able to recognize whether a particular die landed or didn't land on a particular face.} and, consequently, the power set is the $\sigma$-algebra.
-In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold.\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variable which are verifiable up to finite but arbitrarily small precision.}  The probability measure is countably additive, meaning that, given a countable collection $A_i$ of pair-wise disjoint events (i.e.
+In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold.\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variables which are verifiable up to finite but arbitrarily small precision.}  The probability measure is countably additive, meaning that, given a countable collection $A_i$ of pair-wise disjoint events (i.e.
 $A_i \cap A_j = \emptyset$), the probability of the union is the sum of the probability of the respective events:
@@
 \end{equation}
-This last requirement codifies the idea that disjoint event are mutually exclusive: if $A$ and $B$ cannot happen at the same time, then the probability of one of the two happening is the sum of them separately happening.
+This last requirement codifies the idea that disjoint events are mutually exclusive: if $A$ and $B$ cannot happen at the same time, then the probability of one of the two happening is the sum of them separately happening.
 
-\textbf{Expectation and integration.} A random variable is a function $F : X \to \mathbb{R}$ that assigns a value to each state.\footnote{The function $f$ must be measurable, in the sense that the inverse of a Borel should be a Borel set.
-We find that the term is confusing for a physics audience, so it is avoided in the main text.} The number of the side of the die, position, momentum are typical examples.
+\textbf{Expectation and integration.} A random variable is a function $F : X \to \mathbb{R}$ that assigns a value to each state.\footnote{The function $f$ must be measurable, in the sense that the inverse of a Borel set should be a Borel set.
+We find that the term is confusing for a physics audience, so it is avoided in the main text.} The number of the side of the die, position, and momentum are typical examples.
 The expectation for a random variable is given by the integral
@@
 In the discrete case this is simply the counting measure (i.e.
-three possible faces of the die corresponds to three possible states), while in the continuous case this corresponds to the Liouville measure (i.e.
+three possible faces of the die correspond to three possible states), while in the continuous case this corresponds to the Liouville measure (i.e.
 volumes in phase space correspond to the count of states).
-This measure is additive over disjoint region as they represent mutually exclusive events.
+This measure is additive over disjoint regions as they represent mutually exclusive events.
 The requirement is that we can assign non-zero probability only to events that have a non-zero count of states.
@@
 
-\textbf{Choquet theory.} Classical statistical mechanics and probability starts by assuming the existence of a state space $X$ and defining ensembles on top of it.
+\textbf{Choquet theory.} Classical statistical mechanics and probability start by assuming the existence of a state space $X$ and defining ensembles on top of it.
 This cannot be generalized to quantum mechanics and beyond.
@@
 First, we have to define what it means to represent an ensemble with a probability measure.
-Let $E$ be a convex subset of a locally convex space $V$ and $X$ the set of extreme point.
-In classical mechanics, $X$ will correspond to the set of Dirac measures, as these cannot seen as a mixture of other measures.
-In quantum mechanics, $X$ corresponds to the pure states, as these cannot be seen as mixture of other states.
+Let $E$ be a convex subset of a locally convex space $V$ and $X$ the set of extreme points.
+In classical mechanics, $X$ will correspond to the set of Dirac measures, as these cannot be seen as a mixture of other measures.
+In quantum mechanics, $X$ corresponds to the pure states, as these cannot be seen as a mixture of other states.
 In other words, the set of states $X$ is recovered as the set of ensembles with the narrowest spread.
@@
 Therefore, in general, quasi-probability distributions can be understood as affine combinations.
-Remarkably, expectation values of polynomial of position and momentum in symmetric ordering coincide with expectations over the Wigner function.
+Remarkably, expectation values of polynomials of position and momentum in symmetric ordering coincide with expectations over the Wigner function.
 For example
@@
 
-Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measure, the difference between classical probability and quantum quasi-probability does not lie in the use of negative probability.
+Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measures, the difference between classical probability and quantum quasi-probability does not lie in the use of negative probability.
 Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events, and all mutually exclusive events are represented by disjoint regions.
@@
 where $F_\psi$ is an entire analytic function, which cannot vanish on an open set without vanishing everywhere, therefore the support of all Husimi Q functions is the whole space $Z$.
-This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support; moreover disjoint subsets of $Z$ cannot be thought of as mutually exclusive events because we are not going to be able to find two distribution with those subsets as their respective supports.
-For the discrete case, a SIC-POVM representation picks $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot be all orthogonal to each other.
+This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support; moreover disjoint subsets of $Z$ cannot be thought of as mutually exclusive events because we are not going to be able to find two distributions with those subsets as their respective supports.
+For the discrete case, a SIC-POVM representation picks $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot all be orthogonal to each other.
 For a Bloch sphere, the four corners of the tetrahedron are not all opposite to each other.
 Therefore, disjoint sets of $P_i$ do not correspond to mutually exclusive events.
-Quasi-probability representation, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
+Quasi-probability representations, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
 
@@
 \textbf{Ensemble space embedding and signed measures.} We saw that negative probability is a red herring, and mutual exclusivity is key.
-But how exactly mutual exclusivity give us classical probability?
+But how exactly does mutual exclusivity give us classical probability?
 Moreover, are quasi-probability representations specific to quantum mechanics, or are they more general?
@@
 
-In Choquet theory, we saw that classical spaces, Choquet simplexes, are those in which there is a one-to-one correspondence between probability measures and ensembles, where no multiple decomposition in terms of pure state is allowed.
+In Choquet theory, we saw that classical spaces, Choquet simplexes, are those in which there is a one-to-one correspondence between probability measures and ensembles, where no multiple decomposition in terms of pure states is allowed.
 We want to show that this is a consequence of assuming mutual exclusivity between the points of $X$.
@@
 That is, we can move probability from $A$ to $B$ while leaving the ensemble, and therefore the entropy and all statistical variables, unchanged.
-If $A$ and $B$ where two mutually exclusive events, this would not be possible: if we could distinguish between $A$ and $B$, something would have to change.
+If $A$ and $B$ were two mutually exclusive events, this would not be possible: if we could distinguish between $A$ and $B$, something would have to change.
 The only way for $A$ and $B$ to be mutually exclusive, then, is if different measures represent different ensembles, which means the ensemble space is a simplex.
@@
 In quantum mechanics, a classical context is a simplex embedded within the ensemble space.
-For the Bloch ball, a classical context is an axis that connects two opposite point.
+For the Bloch ball, a classical context is an axis that connects two opposite points.
 The points are orthogonal because the midpoint is the maximally mixed state, which therefore maximizes the entropy increase, and the line segment is a one simplex.
-If we take the position operator and consider all density operators that commute with $x$, this is corresponds to the set of functions $\rho(x)$ that integrate two once, which is the space of probability supported by $X$, which is a simplex.
+If we take the position operator and consider all density operators that commute with $x$, this is corresponds to the set of functions $\rho(x)$ that integrate to one, which is the space of probability supported by $X$, which is a simplex.
 
@@
 Wigner functions, for example, are defined over phase space, which does not represent quantum states.
-Let $M^{\pm}(Z)$ the space of all bounded signed measures over $Z$.
-Since the linear combination of two bounded signed measure is a bounded signed measure, $M^{\pm}(Z)$ is a vector space.
+Let $M^{\pm}(Z)$ be the space of all bounded signed measures over $Z$.
+Since the linear combination of two bounded signed measures is a bounded signed measure, $M^{\pm}(Z)$ is a vector space.
 Now let $M_1^{\pm}(Z)$ be the space of signed probability measures.
@@
 Recall that an ensemble space must be a convex subset of a vector space.
-If we are able to represent ensembles as quasi-probability distribution over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such an $Z$ always exists is still open for us.
+If we are able to represent ensembles as quasi-probability distributions over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such an $Z$ always exists is still open for us.
 A possible conjecture is that $Z$ could be constructed from the statistical quantities $F_i$ that fully identify each ensemble in $\Ens$.} In this case, any element $v\in V$ can be represented uniquely by a signed probability measure $\mu_{v}$, regardless of whether it is an actual ensemble or not.
@@
 
-It should be clear that, since an ensemble space is a strict subset of the embedding vector space, not all affine measure will correspond to an ensemble space.
+It should be clear that, since an ensemble space is a strict subset of the embedding vector space, not all affine measures will correspond to an ensemble space.
 Whether all probability measures correspond to a valid ensemble, or whether signed probability measures are even used depends on how the mapping is done in each particular case.
@@
 
-\textbf{Non-uniquess of quasi-probability.} Our approach gives us a consistent view of both quasi-probabilities and classical probability.
-The first are affine representations of the ensembles that connect to the vector space structure, while the second are convex representation of the extreme points which rely on the shape of the ensemble space.
+\textbf{Non-uniqueness of quasi-probability.} Our approach gives us a consistent view of both quasi-probabilities and classical probability.
+The first are affine representations of the ensembles that connect to the vector space structure, while the second are convex representations of the extreme points which rely on the shape of the ensemble space.
 
@@
 However, this practical convenience is also a drawback for finding a more general approach to probability in spaces beyond classical mechanics.
-Crucially, the space $Z$ over which the quasi-probability measures are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tells us which quasi-probability measure represent valid ensembles and which do not.
+Crucially, the space $Z$ over which the quasi-probability measures are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tell us which quasi-probability measures represent valid ensembles and which do not.
 
@@
 Note that the linearity in our axioms is in terms of mixing coefficients, not probability of outcomes.
-That is, regardless of whether we can characterize experiment outcomes with probability distributions, preparation procedures will always allow us to create statistical mixture.
-As we saw in Choquet theory, the difficulty is that there are multiple was to mix pure states to recover the same ensemble.
+That is, regardless of whether we can characterize experiment outcomes with probability distributions, preparation procedures will always allow us to create a statistical mixture.
+As we saw in Choquet theory, the difficulty is that there are multiple ways to mix pure states to recover the same ensemble.
 However, the question of what is the biggest mixture coefficient one can use in a specific case is well-defined, and generalizes to the non-compact case as well.
@@
 Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.
-mixtures),\footnote{Since we work with convex sets within a topology, the hull is both the closed under the topology and convex combinations.
+mixtures),\footnote{Since we work with convex sets within a topology, the hull is both closed under the topology and convex combinations.
 That is, $\hull(A)$ is the smallest closed and convex set that contains $A$} we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
@@
 \textbf{Compare to Choquet theory.} In case $E$ is compact, we can restrict the fraction capacity to the set $X$ of extreme points.
-Given a set $A \subset X$, of extreme points, we can find the set $N \subset E$ of all ensemble that can be represented with a probability measure with support $A$.
+Given a set $A \subset X$, of extreme points, we can find the set $N \subset E$ of all ensembles that can be represented with a probability measure with support $A$.
 This forms a convex set.
@@
 Suppose, in fact, that $p_1$ and $p_2$ are two distinct measures that represent the same ensemble $\ens$.
-Then there will be a set $A$ such such that $p_1(A) \neq p_2(A)$.
+Then there will be a set $A$ such that $p_1(A) \neq p_2(A)$.
 Suppose, without loss of generality, that $p_1(A) > p_2(A)$ We will also have $p_1(A^{\complement}) = 1- p_1(A) < 1 - p_2(A) = p_2(A^{\complement})$.
@@
 Moreover, the setup is more general than Choquet theory (i.e.
-the ensemble space need not be compact) and stems from necessary physical requirement.
+the ensemble space need not be compact) and stems from necessary physical requirements.
 An in depth study on fraction capacities, then, is guaranteed to be applicable to all physical theories, including those yet to be formulated.
 
-One important feature, for example, would be to characterize what fraction capacities are possible for each ensembles space.
+One important feature, for example, would be to characterize what fraction capacities are possible for each ensemble space.
 In the classical case, for example, only additive fraction capacities are possible; in the quantum case, only some type of non-additivity will be possible.
-The shape of the space effectively is encoded in the space of possible fraction capacities, maybe in a way that is mathematically and physically insightful.
+The shape of the space is effectively encoded in the space of possible fraction capacities, maybe in a way that is mathematically and physically insightful.
 
@@
 The entropy volume is a function of the ensemble space, so it is not a measure.
-Let's see everything fits together.
+Let's see how everything fits together.
 
@@
 In classical mechanics, all states are perfectly distinguishable as they represent mutually exclusive cases, the measure is additive.
-In quantum mechanics, only orthogonal state are distinguishable, therefore the measure is non-additive over pure states.
+In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states.
 We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutually exclusiveness, is responsible for the failure of the additivity of the measure.
@@
 The first difficulty, as we already pointed out, is that the ensemble space $\Ens$ may not have extreme points.
-What is the correct procedure to ``close'' the ensembles space and recover the space of extreme points is, for us, an open question.
+What is the correct procedure to ``close'' the ensemble space and recover the space of extreme points is, for us, an open question.
 For now, we will assume $E$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
@@
 
-Recall that a statistical variable is linear function $F : E \to \mathbb{R}$ defined over the ensemble space.
+Recall that a statistical variable is a linear function $F : E \to \mathbb{R}$ defined over the ensemble space.
 Given that $F$ is a linear, its restriction on $X$ will determine the values for all other points of $E$.
@@
 \end{equation}
-This last requirement may not be possible, or it may be used as guide in case that the previous derivative is not uniquely defined.
+This last requirement may not be possible, or it may be used as a guide in case that the previous derivative is not uniquely defined.
 
-\textbf{Ruling out Sugeno and Choquet integrals.} Given that extensive literature already exists for non-additive measure theory (cite), as it is used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence, we originally thought it would be just a matter of clarifying our definitions and connect them to the existing literature.
+\textbf{Ruling out Sugeno and Choquet integrals.} Given that extensive literature already exists for non-additive measure theory (cite), as it is used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence, we originally thought it would be just a matter of clarifying our definitions and connecting them to the existing literature.
 However, things turned out to be more complicated.
-Depending on the use cases, different types of integrals exists.
+Depending on the use cases, different types of integrals exist.
 In fact, procedures exist\cite{nonadditive2014,pap2013} to construct infinitely many types of non-additive integrals.\footnote{Briefly, one can define a pseudo-addition and a pseudo-multiplication as real functions of two real arguments and construct an integral based on those definitions.} At this point, we can rule out the two most common non-additive integrals: the Sugeno integral and the Choquet integral.
@@
 However, it does not work in the quantum case.
-We can show this by computing the Choquet integral for a two states quantum system and see that it does not recover the expectation of the statistical variables in the corresponding ensemble space.
+We can show this by computing the Choquet integral for a two state quantum system and see that it does not recover the expectation of the statistical variables in the corresponding ensemble space.
 
@@
 A statistical variable $F$ (i.e.
-an observables), is a linear function of the ensemble space.
+an observable), is a linear function of the ensemble space.
 This means that two opposite points on the surface $\ens[a]$ and $\ens[b]$ (i.e.
@@
 This means that $A(s)$ will include two opposite points.
-Since $\ens[o]$ can be expressed as the mixture of two opposite point, $\ens[o]$ is in the hull of $A(s)$ and therefore $\frcap_{\ens[o]}(A(s)) = 1$.
+Since $\ens[o]$ can be expressed as the mixture of two opposite points, $\ens[o]$ is in the hull of $A(s)$ and therefore $\frcap_{\ens[o]}(A(s)) = 1$.
 Conversely, if $s > F(\ens[b])$, $A(s) = \emptyset$, which means $\frcap_{\ens[o]} = 0$.
@@
 Conversely, all statistical variables are linear on the ensemble space, so the restriction on the extreme points cannot be arbitrary.
-In the Bloch ball case, for example, all linear functions are fully determined by two opposite points on the sphere and by two value (i.e.
+In the Bloch ball case, for example, all linear functions are fully determined by two opposite points on the sphere and by two values (i.e.
 the eigenstates and the eigenvalues of the corresponding operator).
@@
 In terms of statistical variables, two mutually exclusive events can be assigned values independently.
-This, again, recovers how observable work in quantum mechanics, without mentioning quantum mechanics: only orthogonal states, that is mutually exclusive events, can associated to measurement outcomes and be given a classical probability distribution.
+This, again, recovers how observable work in quantum mechanics, without mentioning quantum mechanics: only orthogonal states, that is mutually exclusive events, can be associated to measurement outcomes and be given a classical probability distribution.
 The non-additivity of the fraction capacity and the non-arbitrariness of the statistical variables over the extreme points, then, must be linked to the shape of the ensemble space.
 
-Additionally, the non-additivity of the state capacity also depends on the degree of mutually exclusiveness of the events.
+Additionally, the non-additivity of the state capacity also depends on the degree of mutual exclusiveness of the events.
 It is additive for mutually exclusive events and sub-additive otherwise.

### 2025-07-28 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Include TikZ code
Replace the PDF with the TikZ code for the Choquet Integral figure.

--- 2025-07-28 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 \usepackage{tikz}
+\usetikzlibrary{3d, perspective}
 \usepackage{breakurl}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{ChoquetCalculation1.jpg}
+
+	% perspective
+	\newcommand{\yaw}{11}
+	\newcommand{\pitch}{28}
+	\newcommand{\roll}{23}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	% dimensions
+	\newcommand{\saxisoffset}{1.5}
+	\newcommand{\canvasaxisoffset}{-3}
+	\newcommand{\s}{-0.4}
+	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
+	\pgfmathsetmacro{\stheta}{asin(\s)}
+	\pgfmathsetmacro{\minstheta}{asin(-\s)}
+
+	\newcommand{\radius}{1}
+	\pgfmathsetmacro{\axislength}{2*\radius}
+	\newcommand{\margin}{1.3}
+
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.3]
+		% Bloch sphere (outline drawn in canvas plane)
+		\draw (0, 0, 0) circle (\radius cm);
+
+		% special points
+		\shadedraw (0, 0, 0) circle (0.03 cm) node[below] {$\mathsf{o}$};
+		\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+		\shadedraw (0, -\radius, 0) circle (0.03cm) node[below] {$\mathsf{a}$};
+
+		% s axis
+		\begin{scope}[shift={(1.95*\xx, 1.95*\yx, 1.95*\zx)}]
+			\draw (0, -1, 0) -- (0, 0, 0);
+			\draw[->] (0, 0, 0) -- (0, 1, 0) node[below left] {$s$};
+		\end{scope}
+
+		% cut
+		\begin{scope}[canvas is zx plane at y=\s]
+			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+		\end{scope}
+
+		% meridians and equator
+		\begin{scope}[canvas is yz plane at x=0]
+			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+			\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is zx plane at y=0]
+			\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+			\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is xy plane at z=0]
+			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+		\end{scope}
+
+		% empty case
+		\draw (3.75*\xx+\xy, 3.75*\yx+\yy, 3.75*\zx+\zy) node {$s > b$};
+
+		% northern case
+		\begin{scope}[shift={(4.5*\xx-0.2*\xy, 4.5*\yx-0.2*\yy, 4.5*\zx-0.2*\zy)}, scale=0.5]
+			\draw (-3.5, +1.3, 0) node {$\frac{1}{2} ( a + b ) < s < b$};
+
+			% north pole
+			\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+
+			% Bloch sphere
+			\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, radius=1 cm];
+			\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, radius=1 cm];
+
+			% cut
+			\begin{scope}[canvas is zx plane at y=-\s]
+				\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+			\end{scope}
+
+			% meridians
+			\begin{scope}[canvas is yz plane at x=0]
+				\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=90-\minstheta, radius=1 cm];
+				\draw[dashed] (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=\minstheta-90, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is xy plane at z=0]
+				\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=180-\minstheta, radius=1 cm];
+				\draw[dashed] (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=\minstheta, radius=1 cm];
+			\end{scope}
+		\end{scope}
+
+		% southern case
+		\begin{scope}[shift={(3*\xx-\xy, 3*\yx-\yy, 3*\zx-\zy)}, scale=0.5]
+			\draw (3.5, -0.5, 0) node {$s < \frac{1}{2} ( a + b )$};
+
+			% north pole
+			\shadedraw (0, 0, 0) circle (0.03 cm) node[below] {$\mathsf{o}$};
+			\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+
+			% Bloch sphere
+			\draw (\xy, \yy, \zy) arc[start angle=90, end angle={-acos(\radiusats*\iyx*\xx + \s*\yx + \radiusats*\iyz*\zx)}, radius=1 cm];
+			\draw (\xy, \yy, \zy) arc[start angle=90, end angle={360-acos(-\radiusats*\iyx*\xx + \s*\yx - \radiusats*\iyz*\zx)}, radius=1 cm];
+			% cut
+			\begin{scope}[canvas is zx plane at y=\s]
+				\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+			\end{scope}
+
+			% meridians and equator
+			\begin{scope}[canvas is yz plane at x=0]
+				\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=90-\stheta, radius=1 cm];
+				\draw[dashed] (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=\stheta-90, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is zx plane at y=0]
+				\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+				\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is xy plane at z=0]
+				\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=180-\stheta, radius=1 cm];
+				\draw[dashed] (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=\stheta, radius=1 cm];
+			\end{scope}
+		\end{scope}
+	\end{tikzpicture}
 	\caption{...}

### 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Minor cleanup
Remove unused parameters
Squashed commit of the following:
commit aa44b1faa3e4bb0281efd2a864bfeccffdd245bd
Author: thrien <39382079+thrien@users.noreply.github.com>
Date:   Mon Jul 28 19:09:00 2025 -0400

    NonAdditiveProbability: Minor cleanup

    Remove unused parameters

commit 61c75fe7e46ce61de3924653124e6ff70989eae6
Author: thrien <39382079+thrien@users.noreply.github.com>
Date:   Mon Jul 28 19:01:23 2025 -0400

    NonAdditiveProbability: Include TikZ code

    Replace the PDF with the TikZ code for the Choquet Integral figure.

commit b7a4b10238c

--- 2025-07-28 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{ChoquetCalculation1.jpg}
+
+	% perspective
+	\newcommand{\yaw}{11}
+	\newcommand{\pitch}{28}
+	\newcommand{\roll}{23}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	% dimensions
+	\newcommand{\s}{-0.4}
+	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
+	\pgfmathsetmacro{\stheta}{asin(\s)}
+	\pgfmathsetmacro{\minstheta}{asin(-\s)}
+
+	\newcommand{\radius}{1}
+	\newcommand{\margin}{1.3}
+
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.3]
+		% Bloch sphere (outline drawn in canvas plane)
+		\draw (0, 0, 0) circle (\radius cm);
+
+		% special points
+		\shadedraw (0, 0, 0) circle (0.03 cm) node[below] {$\mathsf{o}$};
+		\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+		\shadedraw (0, -\radius, 0) circle (0.03cm) node[below] {$\mathsf{a}$};
+
+		% s axis
+		\begin{scope}[shift={(1.95*\xx, 1.95*\yx, 1.95*\zx)}]
+			\draw (0, -1, 0) -- (0, 0, 0);
+			\draw[->] (0, 0, 0) -- (0, 1, 0) node[below left] {$s$};
+		\end{scope}
+
+		% cut
+		\begin{scope}[canvas is zx plane at y=\s]
+			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+		\end{scope}
+
+		% meridians and equator
+		\begin{scope}[canvas is yz plane at x=0]
+			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+			\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is zx plane at y=0]
+			\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+			\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is xy plane at z=0]
+			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+		\end{scope}
+
+		% empty case
+		\draw (3.75*\xx+\xy, 3.75*\yx+\yy, 3.75*\zx+\zy) node {$s > b$};
+
+		% northern case
+		\begin{scope}[shift={(4.5*\xx-0.2*\xy, 4.5*\yx-0.2*\yy, 4.5*\zx-0.2*\zy)}, scale=0.5]
+			\draw (-3.5, +1.3, 0) node {$\frac{1}{2} ( a + b ) < s < b$};
+
+			% north pole
+			\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+
+			% Bloch sphere
+			\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, radius=1 cm];
+			\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, radius=1 cm];
+
+			% cut
+			\begin{scope}[canvas is zx plane at y=-\s]
+				\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+			\end{scope}
+
+			% meridians
+			\begin{scope}[canvas is yz plane at x=0]
+				\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=90-\minstheta, radius=1 cm];
+				\draw[dashed] (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=\minstheta-90, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is xy plane at z=0]
+				\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=180-\minstheta, radius=1 cm];
+				\draw[dashed] (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=\minstheta, radius=1 cm];
+			\end{scope}
+		\end{scope}
+
+		% southern case
+		\begin{scope}[shift={(3*\xx-\xy, 3*\yx-\yy, 3*\zx-\zy)}, scale=0.5]
+			\draw (3.5, -0.5, 0) node {$s < \frac{1}{2} ( a + b )$};
+
+			% north pole
+			\shadedraw (0, 0, 0) circle (0.03 cm) node[below] {$\mathsf{o}$};
+			\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+
+			% Bloch sphere
+			\draw (\xy, \yy, \zy) arc[start angle=90, end angle={-acos(\radiusats*\iyx*\xx + \s*\yx + \radiusats*\iyz*\zx)}, radius=1 cm];
+			\draw (\xy, \yy, \zy) arc[start angle=90, end angle={360-acos(-\radiusats*\iyx*\xx + \s*\yx - \radiusats*\iyz*\zx)}, radius=1 cm];
+			% cut
+			\begin{scope}[canvas is zx plane at y=\s]
+				\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+			\end{scope}
+
+			% meridians and equator
+			\begin{scope}[canvas is yz plane at x=0]
+				\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=90-\stheta, radius=1 cm];
+				\draw[dashed] (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=\stheta-90, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is zx plane at y=0]
+				\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+				\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is xy plane at z=0]
+				\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=180-\stheta, radius=1 cm];
+				\draw[dashed] (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=\stheta, radius=1 cm];
+			\end{scope}
+		\end{scope}
+	\end{tikzpicture}
 	\caption{...}

### 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: adjusting size of figure and caption

@@
 
-	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.3]
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
 		% Bloch sphere (outline drawn in canvas plane)
@@
 	\end{tikzpicture}
-	\caption{...}
+	\caption{For Choquet integral calculation, sets $A(s)$ of all pure states for which the statistical variable is greater than $s$.
+On the left, the Bloch ball with a plane at constant $s$.
+On the right, the top case is when the value is greater than the maximum (empty set); the middle case is when the value is in between the target point $\ens[o]$ and the maximum; the third case is when the value is below the target point.}
 \end{figure}

### 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Label cases in figure

@@
 \usepackage{float}
+\usepackage{subcaption}
 
@@
 In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts, which stems from the underlying aim to characterize physical theories from an operational viewpoint.
-
-in terms of operations aspects, which lends themselves ,  Given the statistical/probabilistic nature of measurement outcomes and the success of information theory, a generalized notion of probability is typically assumed.
-.
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
 A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
+Quasiprobabilistic approaches are also used in quantum mechanics, but it is not completely clear how exactly they relate to standard one.
 
@@
 it is only a mixture of itself).
-
-TODO: figure with classical triangle and quantum Bloch Ball
-
-The set of all extreme points $X$, as we will see, plays a major role.
-However, on physical grounds, these may not be proper ensembles.
-In classical mechanics, these would correspond to Dirac measures: probability distributions all concentrated at a single point of phase space.
-Since we cannot reliably prepare a particle with infinite precision, and its entropy would correspond to minus infinity, these are not ensembles and they are not part of $\Ens$.
-For the purpose of Choquet theory, we will assume that there is a way to find the closure $E \supset \Ens$ that includes all extreme points $X$.
-Moreover, we will also assume $E$ to be compact, even though it will not be in general.
-How to properly make these identifications, and how to go beyond the compact case, is outside of the scope of this particular article.\footnote{An early attempt used the notion of orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
-This does not work.
-An approach, currently under investigation, is to study limits over statistical quantities.
-We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} Note that the classical discrete and the quantum case fall in the compact case.
-Moreover, the set of all probability distributions over a compact space $X$ also falls in the compact case.
-Therefore, apart from the technicality above, the compact case is still a very interesting case.
-To distinguish the context, in this article we will use $\Ens$ to indicate the general case and $E$ to indicate this ``compact closure.''
-
-\textbf{Orthogonality and separatedness.} We will say that two ensembles are orthogonal, noted $\ens[a] \ortho \ens[b]$, when their mixture gives a maximal entropy increase.
-This will recover the notion of orthogonality for both classical and quantum spaces.
-Conceptually, these correspond to two ensembles where all instances are distinct from the other: they are mutually exclusive and a single instance is enough to distinguish between the two cases.
-
-In classical spaces, orthogonality and separatedness coincide: two ensembles are orthogonal and separate if and only if the support of the respective probability distributions overlap on a non-measure-zero set.
-In quantum mechanics, these two notions are different: any two pure states are separate but are not orthogonal.
-For example, pure ensembles of spin up and spin down are orthogonal, mutually exclusive, precisely because it takes one instance to determine which one is which.
-Pure ensembles of spin up and spin right, however, are not orthogonal precisely because, in some cases, their instance cannot be perfectly discriminated.
-They are, however, orthogonal, because they are both pure states, one is not the mixture of other ensembles: they are extreme points.
-
-It is the ease with which such distinctions between the classical and quantum case can be made that is the main appeal for this structure.
-We will see over and over how mutual exclusivity is linked to classical probability and its failure in quantum, and other non-classical, theories.
-
-To summarize, a repeatedly verifiable physical theory requires ensembles, which must be defined experimentally, allow statistical mixing and provide an entropy.
-Additionally, we require the existence of enough statistical quantities to identify each ensemble.
-Under these assumptions, an ensemble space $\Ens$ is a convex set of a metrizable second countable locally convex topological vector space $V$.
-In some cases, for simplicity, we will require $\Ens$ to admit a ``compact closure'' $E$, so that it includes the set $X$ of its extreme points.
-We now turn to the question of how we can represent both the entropy and each ensemble in terms of some measure theoretic structure.
-
-\section{Probability measures and Choquet theory}
-
-\textbf{Classical probability measures.} In the classical case, given the topological space of possible states $X$ (i.e.
-the sample space), ensembles are defined as probability measures $p : \Sigma_X \to [0,1]$.
-That is, a probability measure assigns to each element of the Borel algebra $\Sigma_X$\footnote{In our overall project, Borel sets represent statements associated to a test, regardless of termination.}, an event, the probability to find it to be true.
-In the discrete case, like for roll of a die, $X$ is a set endowed with the discrete topology\footnote{Physically, the discrete topology corresponds to the ability to experimentally verify or falsify each case.
-That is, we are able to recognize whether a particular die landed or didn't land on a particular face.} and, consequently, the power set is the $\sigma$-algebra.
-In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold.\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variables which are verifiable up to finite but arbitrarily small precision.}  The probability measure is countably additive, meaning that, given a countable collection $A_i$ of pair-wise disjoint events (i.e.
-$A_i \cap A_j = \emptyset$), the probability of the union is the sum of the probability of the respective events:
-\begin{equation}
-	p\left(\bigcup_i A_i \right) = \sum_i p(A_i).
-\end{equation}
-This last requirement codifies the idea that disjoint events are mutually exclusive: if $A$ and $B$ cannot happen at the same time, then the probability of one of the two happening is the sum of them separately happening.
-
-\textbf{Expectation and integration.} A random variable is a function $F : X \to \mathbb{R}$ that assigns a value to each state.\footnote{The function $f$ must be measurable, in the sense that the inverse of a Borel set should be a Borel set.
-We find that the term is confusing for a physics audience, so it is avoided in the main text.} The number of the side of the die, position, and momentum are typical examples.
-The expectation for a random variable is given by the integral
-\begin{equation}
-	E[F] = \int_X F dp.
-\end{equation}
-The expectation is linear with respect to the variable (i.e.
-$E[aF+bG]= aE[F]+bE[G]$).
-In our case, what is more interesting is that it is a convex function in terms of probability measures: if $p = \lambda p_1 + \bar{\lambda} p_2$ then $\int_X F dp = \lambda \int_X F dp_1 + \bar{\lambda} \int_X F dp_2$.
-That is, as we required for statistical variables, the expectation for a mixture is the mixture of the expectations.
-
-\textbf{Absolute continuity of probability measures.} As we saw before, probability measures over a single point over the continuum cannot be understood as valid ensembles.
-To address this, we add a requirement that is both physically meaningful and mathematically useful.
-Let $\mu : \Sigma_X \to [0, + \infty]$ be a measure on $X$ that quantifies the count of states in each region.
-In the discrete case this is simply the counting measure (i.e.
-three possible faces of the die correspond to three possible states), while in the continuous case this corresponds to the Liouville measure (i.e.
-volumes in phase space correspond to the count of states).
-This measure is additive over disjoint regions as they represent mutually exclusive events.
-The requirement is that we can assign non-zero probability only to events that have a non-zero count of states.
-That is, if for a particular event $A$ we have $p(A) \neq 0$, then we must have that $\mu(A) \neq 0$ as well.
-Mathematically, $p$ is absolutely continuous with respect to $\mu$, and this is exactly the case in which a probability density $\rho = \frac{dp}{d\mu}$ can be defined.\footnote{In the general case, we can still assign non-zero probability to a point of a manifold, provided that the topology and the measure $\mu$ consistently allow for it.
-Physically, it means that those points are ``special'' both in terms of experimental verifiability and entropy/count of states.
-While the general framework allows for it, it is not crucial for our discussion here.} This is another example in which a physical requirement (probability must be zero if there are no possible states), once properly spelled out (absolute continuity), justifies common physical intuition (we assume a probability density exists).
-
-\textbf{Choquet theory.} Classical statistical mechanics and probability start by assuming the existence of a state space $X$ and defining ensembles on top of it.
-This cannot be generalized to quantum mechanics and beyond.
-In our theory of ensemble spaces we start with ensembles, and have to show when and how they can be thought, can be represented, as probability measures.
-If we can assume that $E$ is a compact convex set, Choquet theory answers this question.
-
-First, we have to define what it means to represent an ensemble with a probability measure.
-Let $E$ be a convex subset of a locally convex space $V$ and $X$ the set of extreme points.
-In classical mechanics, $X$ will correspond to the set of Dirac measures, as these cannot be seen as a mixture of other measures.
-In quantum mechanics, $X$ corresponds to the pure states, as these cannot be seen as a mixture of other states.
-In other words, the set of states $X$ is recovered as the set of ensembles with the narrowest spread.
-We say that $\ens \in E$ is represented by a probability measure $p : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_E f dp$ for every continuous linear functional $f$ on $V$.
-Moreover, $p$ is supported by the extreme points if $p(X) = 1$.
-Physically, we start with an ensemble $\ens$ taken as a point in our ensemble space, identified by the value of all possible statistical variables $f(\ens)$, and we represent it as a probability measure $p$, a statistical mixture, over other the extreme points, the states, recovering the expectation with an integral.
-
-We should note that, given any probability measure $p$ over the extreme points $X$, we can always find an $\ens \in E$ that is represented by $p$.
-Therefore, we have a map from the set of probability measures to the set of ensembles.
-The first question, then, is whether we can go back.
-Choquet theory tells us that, in the compact case, we can:
-\begin{thrm}[Choquet 1]
-	Let $E$ be a metrizable compact convex subset of a locally convex space $V$, and let $\ens \in E$.
-Then there exists a probability measure $p_{\ens}$ on $X$ which represents $\ens$ and is supported by the extreme points $X$.
-\end{thrm}
-Note that an ensemble space is a metrizable convex subset of a locally convex TVS. Therefore, if $E$ is compact, the Choquet theorem applies.
-
-The above theorem tells us that the map between probability measures and ensembles is surjective. The second question, then, when is it bijective? Again, Choquet theory has a clear answer.
-\begin{thrm}[Choquet 2]
-	Let $E$ be a closed convex metrizable subset of a locally convex space $V$.
-Then $E$ is a Choquet simplex if and only if for each $\ens \in E$ there exists a unique measure $p_{\ens}$ which represents $\ens$ and is supported by the extreme points $X$.
-\end{thrm}
-The Choquet simplex is a generalization of the finite dimensional simplex, whose mathematical characterization is rather technical.
-The key point is that $E$ is a Choquet simplex if and only if the ensemble space does not allow multiple decomposition in terms of pure states.
-This confirms the physical intuition from classical and quantum statistical mechanics.
-
-For example, the ensemble space for a two state quantum system, the Bloch ball, is a compact convex subset.
-Any internal point can be seen as the convex combination of points on the surface, the pure states.
-The center, which represents the maximally mixed state, can be thought as an equal mixture of any two opposite points, or a uniform distribution over the whole surface.
-The fact that any ensemble can be seen as a mixture of a set of pure states, then, is reflected by the first part of the Choquet theorem, and it is valid for all physical theories.
-Uniqueness of the representation fails, however, precisely because the ensemble space is not a simplex: it is not a classical space.
-
-The core insight from Chouqet theory, then, is that both in the classical and quantum case we can represent ensembles as a probability distribution over a set of states.
-However, only in the classical case this representation is unique.
-In the quantum case, mixed states can be decomposed into pure states in multiple ways, and it is this multiple decomposition that makes the ensemble space not a simplex, and therefore the space not classical.
-Unique representation of ensembles is a desirable property, and standard probability measures are not going to provide it in general.
-
-\section{Quasiprobability}
-
-%Goal of this section: quasi-probability representations are allowed by the embedding in a topological vector space.
-Choice of representations are guided by what observables are interesting for a specific problem.
-
-%Examples: Wigner, Husimi, Glauber P, SIC-POVM (space not made of spectra of operators), Kirkwood-Dirac (two-time correlation function - observables not necessarily taken at the same time )
-
-%Things to say: some people think about it as weakining axioms (cite papers on negative probability).
-It's more than that.
-Points describe events that are not necessarily mutually exclusive.
-TODO: example of two Wigner function/Husimi Q would be better that have no overlap but they are not orthogonal and/or two functions that have overlap but are orthogonal.
-
-%Thing to say: all these representations are useful because they preserve the linear structure, meaning that convex combinations of ensembles are convex combinations of quasi-probability; expectation values are linear.
-Check, but it should be true.
-
-% 1 describe quasiprobability approaches; go through examples; point out that the quasi-probability can be negative; that one can recover expectations;
-
-\textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions, of which the Wigner function is probably the most widely known.
-The Wigner transform $W : \Ens \to L^1(\mathbb{R}^2, \mathbb{R}) $, defined as \cite[eq.~11.8-3a]{mandel1995}
-\begin{equation}
-	W(\rho) = W_{\rho} (q, p) = 2 \int_{-\infty}^{\infty} \langle x + y | \rho | x - y \rangle e^{-2 i p y / \hbar} \; \mathrm dy,
-\end{equation}
-takes a mixed state $\rho$ and returns a real function $W_{\rho}$ of phase space that, like a probability density integrates to one, but, unlike a probability density, can be negative.
-Recall that an affine combination is a linear combination where all coefficients sum to one but are not necessarily non-negative.
-Therefore, in general, quasi-probability distributions can be understood as affine combinations.
-Remarkably, expectation values of polynomials of position and momentum in symmetric ordering coincide with expectations over the Wigner function.
-For example
-\begin{equation}
-	\tr \left( \frac{1}{2}[XP+PX] \rho \right) = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} xp \; W_{\rho} (q, p) dq dp \; .
-\end{equation}
-Moreover, the transform is an affine map: $W(p \rho_1 + \bar{p} \rho_2 ) = p W(\rho_1) + \bar{p} W(\rho_2)$.
-
-In some problems, especially in quantum optics, normal ordering and anti-normal ordering are more useful.
-The Glauber P representation and the Husimi Q representation work like the Wigner function except that the expectations of polynomials recover the respective order.
-In other circumstances, quasi-probability distributions over observables different from position and momentum are more useful.
-For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times \cite{lostaglio2023}.
-Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
-For example, the SIC-POVM represents quantum states as affine combinations of pure states $P_i$ that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
-Each ensemble $\rho$ can then be represented as \cite[eqs.
-28-29]{fuchs2004}
-\begin{equation}
-	\rho = \sum_{n=1}^{d^2} \lambda_i P_i = \sum_{n=1}^{d^2} \left[ (d^2 + 1) p_i - \frac{1}{d} \right] P_i \; ,
-\end{equation}
-where $p_i = \tr \left(P_i \rho\right)$.
-
-It should be clear that there are many ways to represent quantum states via quasi-probability distributions, and that the choice of representation is a matter of convenience for the problem at hand.
-All representations, though, have the following features in common.
-First, we have a map $W : \Ens \to Q(Z,\mathbb{R})$ that given each mixed state $\rho \in \Ens$ returns the corresponding quasi-probability function $w(z) \in Q(Z,\mathbb{R})$, where $Q(Z,\mathbb{R}) \subset L^1(Z,\mathbb{R})$ is the function space used by the specific representation.
-The space $Z$, in fact, can change from representation to representation: it is $\mathbb{R}^{2n}$ for a Wigner function, $\mathbb{C}^n$ for the Husimi Q and $[1, d^2] \subset \mathbb{N}$ for the SIC-POVM.
-All representations preserves convex combinations: $W(\sum_i p_i \rho_i) = \sum_i p_i W(\rho_i)$.
-That is, a statistical mixture of quasi-probability distributions represents the statistical mixture of the corresponding states.
-Since the map is linear, and expectation values are linear, one can express expectation values as integrals or sums over the quasi-probability $w(z)$.
-
-% 2 Note that it's not about negative probability; disjoint not mutually exclusive
-
-\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasi-probability \cite{adhikary2020}.
-Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
-This is a mischaracterization.
-For example, the Husimi Q representation provides us with a non-negative quasi-probability function.
-That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.\footnote{Since the Husimi Q is defined as $Q_{\rho} ( \alpha ) = \frac{1}{\pi} \langle \alpha | \rho | \alpha \rangle$, it is proportional to the probability of measuring a coherent (i.e.
-Gaussian state) $| \alpha \rangle$, which is non-negative.
-} Similarly, a SIC-POVM characterizes the Bloch ball by four pure states that form a tetrahedron inscribed in the ball.
-Since we can also inscribe the Bloch ball inside a tetrahedron, we can represent all states of a two-state quantum system by a convex combination of the vertices, leading to a non-negative quasi-probability distribution over four elements.
-
-Conversely, we can use signed measures to represent classical distributions.
-Suppose we have a two state classical discrete space, which means the space of all possible ensembles is spanned by $[p, (1-p)]$.
-Alternatively, we can write it as $p[1,0] + (1-p)[0,1]$ which highlights that we are writing each possible probability distribution in terms of the two possible cases.
-But we could also express each probability distribution in terms of other two distributions, as in $\left(2p - \frac{1}{2}\right)\left[\frac{3}{4}, \frac{1}{4}\right] + \left(\frac{3}{2} - 2p\right)\left[\frac{1}{4}, \frac{3}{4}\right]$.
-The coefficients now span between $\left[-\frac{1}{2}, \frac{3}{2}\right]$, can be negative, but still sum to one.
-That is, in terms of this new affine basis, we need to use a signed probability.
-
-Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measures, the difference between classical probability and quantum quasi-probability does not lie in the use of negative probability.
-Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events, and all mutually exclusive events are represented by disjoint regions.
-Mutually exclusive events can be prepared and observed independently from the others (i.e.
-they are orthogonal in the sense that they maximize entropy increase during mixture); we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasi-probability in quantum mechanics.
-
-For example, given a pure state $|\psi\>$, the Husimi Q function representing it is given by $Q_\psi(\alpha) = \frac{1}{\pi} |\langle \alpha | \psi \rangle|^2$.
-The inner product is given by \cite[eq.
-11.7-1]{mandel1995}
-\begin{equation}
-	\langle \alpha | \psi \rangle = e^{-|\alpha|^2/2} F_\psi(\alpha^\ast)
-\end{equation}
-where $F_\psi$ is an entire analytic function, which cannot vanish on an open set without vanishing everywhere, therefore the support of all Husimi Q functions is the whole space $Z$.
-This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support; moreover disjoint subsets of $Z$ cannot be thought of as mutually exclusive events because we are not going to be able to find two distributions with those subsets as their respective supports.
-For the discrete case, a SIC-POVM representation picks $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot all be orthogonal to each other.
-For a Bloch sphere, the four corners of the tetrahedron are not all opposite to each other.
-Therefore, disjoint sets of $P_i$ do not correspond to mutually exclusive events.
-Quasi-probability representations, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
-
-% 3 These representations are possible because of the vector space structure of the ensemble space, and therefore something similar will be possible in ANY physical theory
-
-\textbf{Ensemble space embedding and signed measures.} We saw that negative probability is a red herring, and mutual exclusivity is key.
-But how exactly does mutual exclusivity give us classical probability?
-Moreover, are quasi-probability representations specific to quantum mechanics, or are they more general?
-This is where the perspective from the ensemble space framework allows us to answer these questions directly.
-
-In Choquet theory, we saw that classical spaces, Choquet simplexes, are those in which there is a one-to-one correspondence between probability measures and ensembles, where no multiple decomposition in terms of pure states is allowed.
-We want to show that this is a consequence of assuming mutual exclusivity between the points of $X$.
-Suppose we have an ensemble $\ens$ that can be represented as different measures $p_1$ and $p_2$ over the extreme points.
-This means that there are two disjoint sets $A, B \subseteq X$ for which $p_1(A) < p_2(A)$ and $p_1(B) > p_2(B)$.
-That is, we can move probability from $A$ to $B$ while leaving the ensemble, and therefore the entropy and all statistical variables, unchanged.
-If $A$ and $B$ were two mutually exclusive events, this would not be possible: if we could distinguish between $A$ and $B$, something would have to change.
-The only way for $A$ and $B$ to be mutually exclusive, then, is if different measures represent different ensembles, which means the ensemble space is a simplex.
-This tells us that we can use classical probability exactly when all extreme points are mutually exclusive (i.e.
-orthogonal as defined from the entropy) and the ensemble space is a simplex.
-
-That is, classical probability does not merely correspond to the use of a single non-negative probability measure.
-It corresponds to the ability to, conceptually, allow any possible probability measure to describe a possible distinct scenario.
-This requirement is exactly mutual exclusivity.
-In quantum mechanics, a classical context is a simplex embedded within the ensemble space.
-For the Bloch ball, a classical context is an axis that connects two opposite points.
-The points are orthogonal because the midpoint is the maximally mixed state, which therefore maximizes the entropy increase, and the line segment is a one simplex.
-If we take the position operator and consider all density operators that commute with $x$, this is corresponds to the set of functions $\rho(x)$ that integrate to one, which is the space of probability supported by $X$, which is a simplex.
-
-Now the second question: why does quantum mechanics allow quasi-probability representations?
-Let $Z$ be the space over which we are defining our quasi-probability distributions.
-Note that, in general, $Z$ is not the space of states.
-Wigner functions, for example, are defined over phase space, which does not represent quantum states.
-Let $M^{\pm}(Z)$ be the space of all bounded signed measures over $Z$.
-Since the linear combination of two bounded signed measures is a bounded signed measure, $M^{\pm}(Z)$ is a vector space.
-Now let $M_1^{\pm}(Z)$ be the space of signed probability measures.
-While it is not a subspace of $M(Z)$, it does not include the zero measure, it is an affine subspace therefore, up to the choice of an arbitrary origin, it is a vector space.
-
-Recall that an ensemble space must be a convex subset of a vector space.
-If we are able to represent ensembles as quasi-probability distributions over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such an $Z$ always exists is still open for us.
-A possible conjecture is that $Z$ could be constructed from the statistical quantities $F_i$ that fully identify each ensemble in $\Ens$.} In this case, any element $v\in V$ can be represented uniquely by a signed probability measure $\mu_{v}$, regardless of whether it is an actual ensemble or not.
-Since statistical quantities $F_i$ are linear functional, we will have $F_i(v) = \int_Z F_i(z) d\mu_{v}$.
-In other words, there is nothing special about quantum mechanics that allows quasi-probability representations.
-It is the fact that the ensemble space embeds into a vector space, a general fact for all theories, that allows such representations.
-
-It should be clear that, since an ensemble space is a strict subset of the embedding vector space, not all affine measures will correspond to an ensemble space.
-Whether all probability measures correspond to a valid ensemble, or whether signed probability measures are even used depends on how the mapping is done in each particular case.
-For example, there is no simple characterization of which Wigner functions correspond to valid quantum states and which do not.
-
-% 4 pros and cons - can choose a repretation for the problem at hand; no fundamental representation; boundaries ore not well defined (i.e. which functions represnt valid states)
-
-\textbf{Non-uniqueness of quasi-probability.} Our approach gives us a consistent view of both quasi-probabilities and classical probability.
-The first are affine representations of the ensembles that connect to the vector space structure, while the second are convex representations of the extreme points which rely on the shape of the ensemble space.
-
-Quasi-probability distributions are very useful in practice, particularly because one can construct them to fit particular problems.
-Once one has identified a set of observables that are the most natural ones to define a problem, one may find a representation that is specifically suited for the case.
-While the quasi-probability cannot be interpreted as probability, it still provides a useful tool to describe and understand the problem and give some intuition.
-
-However, this practical convenience is also a drawback for finding a more general approach to probability in spaces beyond classical mechanics.
-Crucially, the space $Z$ over which the quasi-probability measures are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tell us which quasi-probability measures represent valid ensembles and which do not.
-
-As a comparison, in the previous case all probability measures were given over a uniquely defined set of points, the extreme points of the ensemble space, but the probability measure may not be unique.
-In this case, the quasi-probability associated to each ensemble is unique, but the space over which the quasi-probability is defined is not.
-
-\section{Fraction capacity}
-
-\textbf{Probability of outcomes vs fraction of components.} We saw that classical probability fails in quantum mechanics because pure states are not always mutually exclusive, are not orthogonal.
-That is why, in quantum mechanics, probability can be defined only once a context, an orthogonal basis, is chosen.
-Therefore, at a conceptual level, probabilities can only give us a patchwork representation of ensembles.
-Some author embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.
-[CITATION NEEDED] However, there is a different route.
-
-Note that the linearity in our axioms is in terms of mixing coefficients, not probability of outcomes.
-That is, regardless of whether we can characterize experiment outcomes with probability distributions, preparation procedures will always allow us to create a statistical mixture.
-As we saw in Choquet theory, the difficulty is that there are multiple ways to mix pure states to recover the same ensemble.
-However, the question of what is the biggest mixture coefficient one can use in a specific case is well-defined, and generalizes to the non-compact case as well.
-Therefore, our generalization of probability is not in terms of what an ensemble can lead to (i.e.
-measurements), but in terms of how ensembles can be produced (i.e.
-preparations).
-
-\textbf{Fraction capacity.} Let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
-How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
-For example, let $\Ens$ be the space of all probability distributions for a six-sided die.
-Let $\ens_{123456}$ be the uniform distribution over all outcomes.
-Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
-Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$, where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/6$ of $\ens$ can be constructed from $A$, but no more.
-Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
-Note that, given the uniform distribution, $1/6$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
-This gives the basic insight for our definitions.
-
-Given a target ensemble $\ens \in \Ens$ and an arbitrary ensemble $\ens[a] \in \Ens$, we define the \textbf{fraction} of $\ens[a]$ in $\ens$ to be
-\begin{equation}
-	\fraction_{\ens}(\ens[a]) = \sup(\{ p \in [0,1] \, | \, \exists \, \ens_1 \in \Ens \text{ s.t. }  \ens = p \ens[a] + \bar{p} \ens_1 \}).
-\end{equation}
-The fraction is always well-defined because we can always write $\ens = 0 \ens[a] + 1 \ens$, therefore it must be zero or greater.
-This quantity tells us how much of ensemble $\ens$ can be constructed from $\ens[a]$.
-
-We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a Borel set of ensembles $A \subset \Ens$.
-Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.
-mixtures),\footnote{Since we work with convex sets within a topology, the hull is both closed under the topology and convex combinations.
-That is, $\hull(A)$ is the smallest closed and convex set that contains $A$} we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
-\begin{equation}
-	\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\}).
-\end{equation}
-This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$.
-
-\textbf{Non-additivity of the fraction capacity.} One can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
-\begin{enumerate}
-	\item non-negative and unit bounded - $\frcap_{\ens}(A) \in [0,1]$
-	\item monotone - $A \subseteq B \implies \frcap_{\ens}(A) \leq \frcap_{\ens}(B)$
-	\item sub-additive - $\frcap_{\ens}(A \cup B) \leq \frcap_{\ens}(A) + \frcap_{\ens}(B)$
-	\item continuous from below and above - $\frcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \frcap_{\ens}(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
-\end{enumerate}
-
-The first property readily comes from the range of the fraction.
-For the second, note that $\hull$ and $\sup$ are both monotone.
-For sub-additivity, note that the fraction of two components can at most sum during mixing.
-For the last, increasing and decreasing sequences strictly add or remove possible mixtures, therefore they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
-The limit of these sequences must agree with the fraction capacity of the limit.
-
-Note that, if not for the lack of additivity over disjoint sets, the fraction capacity would be a probability measure.
-It is therefore a non-additive probability measure.
-The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
-fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\footnote{ \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} }
-
-\textbf{Compare to Choquet theory.} In case $E$ is compact, we can restrict the fraction capacity to the set $X$ of extreme points.
-Given a set $A \subset X$, of extreme points, we can find the set $N \subset E$ of all ensembles that can be represented with a probability measure with support $A$.
-This forms a convex set.
-Now, the ability to express an ensemble $\ens$ as the convex combination of $N$ is precisely the ability to find a measure $p$ that represents $\ens$ such that $p(A) \neq 0$.
-That is, $\ens = \lambda \ens[a] + \bar{\lambda} \ens[b]$ with $\ens[a] \in M$ if and only if there is a probability measure $p$ that represents $\ens$ such that $p(A) \geq \lambda$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{p(A) \, | \, p \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all possible measures.
-
-As we saw, the classical case corresponds to a simplex in Choquet theory.
-This means that each element of the ensemble space is represented by a unique probability measure over the extreme points.
-In that case, the fraction capacity restricted to the extreme points will also be additive.
-In the non-classical case, the closure of the ensemble space will not be a simplex, and there will be multiple representations, which will lead to a non-additive measure.
-Suppose, in fact, that $p_1$ and $p_2$ are two distinct measures that represent the same ensemble $\ens$.
-Then there will be a set $A$ such that $p_1(A) \neq p_2(A)$.
-Suppose, without loss of generality, that $p_1(A) > p_2(A)$ We will also have $p_1(A^{\complement}) = 1- p_1(A) < 1 - p_2(A) = p_2(A^{\complement})$.
-Since the fraction capacity for a set cannot be less than the measure on that set, we have $\frcap_{\ens}(A \cup A^{\complement})=\frcap_{\ens}(X) = 1 = p_1(A) + p_1(A^{\complement}) < (A) + p_2(A^{\complement}) \leq \frcap_{\ens}(A) + \frcap_{\ens}(A^{\complement})$.
-That is, the fraction capacity is additive over disjoint sets of extreme points in and only in the classical case.
-In other words, the fraction capacity reduces to a classical probability distribution over pure states in the classical case.
-
-\textbf{Uniqueness of the fraction capacity.} From a conceptual standpoint, the fraction capacity has a major advantage: it has a clear physical meaning and it is unique.
-Like in Choquet theory and unlike for quasi-probability, the domain over which the measure is defined is fully determined by the system, by the ensemble space.
-On the other hand, unlike Choquet theory and like quasi-probability, once the space is fixed, there is one unique fraction capacity for each ensemble.
-Moreover, the setup is more general than Choquet theory (i.e.
-the ensemble space need not be compact) and stems from necessary physical requirements.
-An in depth study on fraction capacities, then, is guaranteed to be applicable to all physical theories, including those yet to be formulated.
-
-One important feature, for example, would be to characterize what fraction capacities are possible for each ensemble space.
-In the classical case, for example, only additive fraction capacities are possible; in the quantum case, only some type of non-additivity will be possible.
-The shape of the space is effectively encoded in the space of possible fraction capacities, maybe in a way that is mathematically and physically insightful.
-
-\section{State capacity}
-
-\textbf{Counting states.} Since the fraction capacity can be seen as the non-additive generalization of a probability measure, we now want to introduce a similar generalization for the notion of count of states.
-Since entropy characterizes the variability of the preparations of the ensemble, greater variability means the ensemble is spread over more cases.
-Therefore the count of distinguishable states must be a monotonic function of the entropy.
-Given that in statistical mechanics the entropy is the logarithm of the count of states, we expect the count of states to be the exponential of the entropy.
-Therefore one can argue that the proper way to measure the spread of an ensemble is through the exponential of the entropy, also known as entropic volume.\footnote{It has already been shown in \cite{hall1999} that the exponential of the entropy is the only function that can be used to define a volume.} However, the count of states, like the probability measure, is typically defined on the sample space $\Omega$, which can only work in the classical case.
-The entropy volume is a function of the ensemble space, so it is not a measure.
-Let's see how everything fits together.
-
-To build an intuition, consider the classical case.
-For every region $U$ of phase space, we have a set of possible probability distributions that have support $U$.
-This is a convex set since every mixture of distributions with support $U$ has support $U$.
-Note that the uniform distribution $\rho_U$, if it exists, is the one with the greatest entropy, which will correspond to $S(\rho_U) = \log \mu(U)$.
-If $U$ is not compact, then there is no upper limit to the entropy and $\mu(U) = +\infty$.
-Now, let us transfer this intuition on the ensemble space $\ens$.
-If we take a set $A$, we can ask: what is the highest entropy reachable by a mixture of elements of $A$?
-If $A$ happens to be the set of ensembles with support $U$, then the supremum of the entropy is $\log \mu(U)$, which works also in the unbounded case.
-
-The above intuition started with the notion of the state space $X$ and the count of states $\mu$.
-In the general case, we will start with the ensemble space $\Ens$ and the entropy $S$.
-For ensembles, we can define the entropic volume\footnote{CITE Hall} to be the exponential of the entropy $2^{S(\ens)}$.
-For the entropic volume to quantify the spread of the ensemble, we must have that the spread can at most sum.
-Remarkably, the bounds on the entropy are such that the exponential of the entropy is subadditive.
-That is, $2^{S(p\ens[a]+\bar{p}\ens[b])} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])}$ for any ensemble space.\footnote{It is an open question whether the exponential is the only monotonic function of the entropy that is subadditive.} For the equality to hold, the ensemble must be orthogonal, meaning that two states are distinguishable if and only if they are mutually exclusive.
-Therefore, this is indeed a general way to define the spread of ensembles and the count of distinguishable states in a way that is consistent with the previous discussions.
-
-\textbf{State capacity.} Given a Borel set of ensembles $A \subseteq \Ens$, we define the \textbf{state capacity} of $A$ to be
-\begin{equation}
-\capacity(A) = \sup(2^{S(\hull(A))}\cup\{0\}).
-\end{equation}
-This returns the upper bound to the exponential of entropy achievable with a mixture of elements of $A$.
-
-One can then show that the fraction capacity $\capacity : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
-\begin{enumerate}
-	\item non-negative - $\capacity(A)\geq 0$
-	\item monotone - $A \subseteq B \implies \capacity(A) \leq \capacity(B)$
-	\item sub-additive - $\capacity(A \cup B) \leq \capacity(A) + \capacity(B)$
-	\item additive over orthogonal sets - $\capacity(A \cup B) = \capacity(A) + \capacity(B)$ if $A \ortho B$
-	\item continuous from below and above - $\capacity(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \capacity(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
-\end{enumerate}
-
-The first property readily comes from the range of the exponential.
-For the second, note that $\hull$, exponential and $\sup$ are all monotone.
-For sub-additivity, the exponential of the entropy itself is sub-additive, and it is additive over orthogonal ensembles.
-For the last, since the entropy itself  is continuous, this property carries over to the set function.
-
-Again we note that the non-additive measure we recover is very well-behaved, and connects directly with important quantities.
-In the compact case, state capacity can be extended to the extreme points.
-In the classical case, disjoint sets of extreme points will be orthogonal, meaning that the state capacity there restricted is an additive measure.
-In the discrete case, every point is associated to zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
-In the continuous case, gives a subset of phase space $U$ and $A_U$ the set of distributions with support within $U$, the maximum entropy is reached by the uniform distribution $\rho_U \in A_U$, which means we recover $\log_2 \capacity(A_U) = S(\rho_U)$, the fundamental postulate of statistical mechanics.
-Similarly, in quantum mechanics, given an orthogonal subspace $A$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(A) = 2^{\dim(A)}$.
-
-The state capacity, then, recovers the number of distinguishable states.
-In classical mechanics, all states are perfectly distinguishable as they represent mutually exclusive cases, the measure is additive.
-In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states.
-We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutually exclusiveness, is responsible for the failure of the additivity of the measure.
-Note that, since the fraction capacity depends only on the ensemble space, and not also on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
-
-\section{Non-additive generalization of probability theory}
-
-\textbf{Requirements for non-additive probability theory.} In a previous section we reviewed what types of objects we have in probability theory for classical ensembles: a sample space $X$, a probability measure $p$, random variables $F : X \to \mathbb{R}$, expectations as integrals of the probability measure $E[F] = \int_X F(x) dp$, a state counting measure $\mu$, the probability density as the Radon-Nikodym derivative $\frac{dp}{d\mu}$ and the entropy as $\int_X \rho \log \rho d\mu$.
-Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the states counting measure $\mu$, is there a generalization of all other elements?
-That is, can we find an integral and derivative operation over non-additive measures that generalizes to all physical theories?
-
-The first difficulty, as we already pointed out, is that the ensemble space $\Ens$ may not have extreme points.
-What is the correct procedure to ``close'' the ensemble space and recover the space of extreme points is, for us, an open question.
-For now, we will assume $E$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
-Therefore we can talk about the restriction on $X$ of the state capacity and the fraction capacity.
-
-Recall that a statistical variable is a linear function $F : E \to \mathbb{R}$ defined over the ensemble space.
-Given that $F$ is a linear, its restriction on $X$ will determine the values for all other points of $E$.
-Since the expectation is recovered through the integration of the variable over the probability density, we would like a non-additive integral $(E)\int_X$ such that
-\begin{equation}
-	E[F] = F(\ens) = (E)\int_X F(x) d\frcap_{\ens}.
-\end{equation}
-which reduces to the Riemann integral in the classical (additive) case.
-The integral will depend only on the values of $F$ and $\frcap_{\ens}$ on $X$, but since these uniquely determine both $F$ and $\ens$, the integral can be understood as a map from $F$ and $\ens$ to $F(\ens)$, which is, in principle, well-defined.
-Therefore, in principle, this integral should exist.
-
-Additionally, we would like to be able to define an analogue of the probability density.
-We are therefore looking for a non-additive derivative $\rho = \frac{d \frcap_{\ens}}{d \capacity}$ that reduces to the Radon-Nikodym derivative in the classical (additive) case such that
-\begin{equation}
-	F(\ens) = (E)\int_X F(x) \rho(x) d\capacity
-\end{equation}
-under the previous integral. It is less clear whether this derivative is always well defined. Ideally, we would also expect
-\begin{equation}
-	S(\ens) = (E)\int_X \rho(x) \log \rho(x) d\capacity.
-\end{equation}
-This last requirement may not be possible, or it may be used as a guide in case that the previous derivative is not uniquely defined.
-
-\textbf{Ruling out Sugeno and Choquet integrals.} Given that extensive literature already exists for non-additive measure theory (cite), as it is used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence, we originally thought it would be just a matter of clarifying our definitions and connecting them to the existing literature.
-However, things turned out to be more complicated.
-Depending on the use cases, different types of integrals exist.
-In fact, procedures exist\cite{nonadditive2014,pap2013} to construct infinitely many types of non-additive integrals.\footnote{Briefly, one can define a pseudo-addition and a pseudo-multiplication as real functions of two real arguments and construct an integral based on those definitions.} At this point, we can rule out the two most common non-additive integrals: the Sugeno integral and the Choquet integral.
-
-The Sugeno integral does not recover the Lebesgue integral in the case the capacity is additive (i.e.
-classical), and is therefore quickly ruled out.
-The Choquet integral does, so it will work in the classical case.
-However, it does not work in the quantum case.
-We can show this by computing the Choquet integral for a two state quantum system and see that it does not recover the expectation of the statistical variables in the corresponding ensemble space.
 
@@
 	% dimensions
-	\newcommand{\s}{-0.4}
+	\newcommand{\radius}{1}
+
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
+		% Triangle
+		\draw (canvas cs: x=-3cm, y=-0.25cm) +(canvas polar cs: angle=90, radius=1cm) -- +(canvas polar cs: angle=210, radius=1cm) -- +(canvas polar cs: angle=330, radius=1cm) -- cycle;
+
+		% Bloch sphere (outline drawn in canvas plane)
+		\draw (0, 0, 0) circle (\radius cm);
+
+		% meridians and equator
+		\begin{scope}[canvas is yz plane at x=0]
+			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+			\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is zx plane at y=0]
+			\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+			\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is xy plane at z=0]
+			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+		\end{scope}
+	\end{tikzpicture}
+	\caption{Examples of a classical discrete ensemble space with three pure states (triangle) and of a two dimensional quantum ensemble space (ball).}
+\end{figure}
+
+The set of all extreme points $X$, as we will see, plays a major role.
+However, on physical grounds, these may not be proper ensembles.
+In classical mechanics, these would correspond to Dirac measures: probability distributions all concentrated at a single point of phase space.
+Since we cannot reliably prepare a particle with infinite precision, and its entropy would correspond to minus infinity, these are not ensembles and they are not part of $\Ens$.
+For the purpose of Choquet theory, we will assume that there is a way to find the closure $E \supset \Ens$ that includes all extreme points $X$.
+Moreover, we will also assume $E$ to be compact, even though it will not be in general.
+How to properly make these identifications, and how to go beyond the compact case, is outside the scope of this particular article.\footnote{An early attempt used the notion of orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
+This does not work.
+An approach, currently under investigation, is to study limits over statistical quantities.
+We are still trying to understand what mathematical structures have enough information to allow the reconstruction.} Note that the classical discrete and the quantum case fall in the compact case.
+Moreover, the set of all probability distributions over a compact space $X$ also falls in the compact case.
+Therefore, apart from the technicality above, the compact case is still a very interesting case.
+To distinguish the context, in this article we will use $\Ens$ to indicate the general case and $E$ to indicate this ``compact closure.''
+
+\textbf{Orthogonality and separatedness.} We will say that two ensembles are orthogonal, noted $\ens[a] \ortho \ens[b]$, when their mixture gives a maximal entropy increase.
+This will recover the notion of orthogonality for both classical and quantum spaces.
+Conceptually, these correspond to two ensembles where all instances are distinct from the other: they are mutually exclusive and a single instance is enough to distinguish between the two cases.
+
+In classical spaces, orthogonality and separatedness coincide: two ensembles are orthogonal and separate if and only if the support of the respective probability distributions overlap on a non-measure-zero set.
+In quantum mechanics, these two notions are different: any two pure states are separate but are not orthogonal.
+For example, pure ensembles of spin up and spin down are orthogonal, mutually exclusive, precisely because it takes one instance to determine which one is which.
+Pure ensembles of spin up and spin right, however, are not orthogonal precisely because, in some cases, their instance cannot be perfectly discriminated.
+They are, however, orthogonal, because they are both pure states, one is not the mixture of other ensembles: they are extreme points.
+
+It is the ease with which such distinctions between the classical and quantum case can be made that is the main appeal of this structure.
+We will see over and over how mutual exclusivity is linked to classical probability and its failure in quantum, and other non-classical, theories.
+
+To summarize, a repeatedly verifiable physical theory requires ensembles, which must be defined experimentally, allow statistical mixing and provide an entropy.
+Additionally, we require the existence of enough statistical quantities to identify each ensemble.
+Under these assumptions, an ensemble space $\Ens$ is a convex set of a metrizable second countable locally convex topological vector space $V$.
+In some cases, for simplicity, we will require $\Ens$ to admit a ``compact closure'' $E$, so that it includes the set $X$ of its extreme points.
+We now turn to the question of how we can represent both the entropy and each ensemble in terms of some measure theoretic structure.
+
+\section{Probability measures and Choquet theory}
+
+\textbf{Classical probability measures.} In the classical case, given the topological space of possible states $X$ (i.e.
+the sample space), ensembles are defined as probability measures $p : \Sigma_X \to [0,1]$.
+That is, a probability measure assigns to each element of the Borel algebra $\Sigma_X$\footnote{In our overall project, Borel sets represent statements associated with a test, regardless of termination.}, an event, the probability to find it to be true.
+In the discrete case, like for rolling a die, $X$ is a set endowed with the discrete topology\footnote{Physically, the discrete topology corresponds to the ability to experimentally verify or falsify each case.
+That is, we are able to recognize whether a particular die landed or didn't land on a particular face.} and, consequently, the power set is the $\sigma$-algebra.
+In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold.\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variables which are verifiable up to finite but arbitrarily small precision.}  The probability measure is countably additive, meaning that, given a countable collection $A_i$ of pair-wise disjoint events (i.e.
+$A_i \cap A_j = \emptyset$), the probability of the union is the sum of the probability of the respective events:
+\begin{equation}
+	p\left(\bigcup_i A_i \right) = \sum_i p(A_i).
+\end{equation}
+This last requirement codifies the idea that disjoint events are mutually exclusive: if $A$ and $B$ cannot happen at the same time, then the probability of one of the two happening is the sum of them separately happening.
+
+\textbf{Expectation and integration.} A random variable is a function $F : X \to \mathbb{R}$ that assigns a value to each state.\footnote{The function $f$ must be measurable, in the sense that the inverse of a Borel set should be a Borel set.
+We find that the term is confusing for a physics audience, so it is avoided in the main text.} The number of the side of the die, position, and momentum are typical examples.
+The expectation for a random variable is given by the integral
+\begin{equation}
+	E[F] = \int_X F dp.
+\end{equation}
+The expectation is linear with respect to the variable (i.e.
+$E[aF+bG]= aE[F]+bE[G]$).
+In our case, what is more interesting is that it is a convex function in terms of probability measures: if $p = \lambda p_1 + \bar{\lambda} p_2$ then $\int_X F dp = \lambda \int_X F dp_1 + \bar{\lambda} \int_X F dp_2$.
+That is, as we required for statistical variables, the expectation for a mixture is the mixture of the expectations.
+
+\textbf{Absolute continuity of probability measures.} As we saw before, probability measures over a single point on the continuum cannot be understood as valid ensembles.
+To address this, we add a requirement that is both physically meaningful and mathematically useful.
+Let $\mu : \Sigma_X \to [0, + \infty]$ be a measure on $X$ that quantifies the count of states in each region.
+In the discrete case this is simply the counting measure (i.e.
+three possible faces of the die correspond to three possible states), while in the continuous case this corresponds to the Liouville measure (i.e.
+volumes in phase space correspond to the count of states).
+This measure is additive over disjoint regions as they represent mutually exclusive events.
+The requirement is that we can assign a non-zero probability only to events that have a non-zero count of states.
+That is, if for a particular event $A$ we have $p(A) \neq 0$, then we must have that $\mu(A) \neq 0$ as well.
+Mathematically, $p$ is absolutely continuous with respect to $\mu$, and this is exactly the case in which a probability density $\rho = \frac{dp}{d\mu}$ can be defined.\footnote{In the general case, we can still assign non-zero probability to a point of a manifold, provided that the topology and the measure $\mu$ consistently allow for it.
+Physically, it means that those points are ``special'' both in terms of experimental verifiability and entropy/count of states.
+While the general framework allows for it, it is not crucial for our discussion here.} This is another example in which a physical requirement (probability must be zero if there are no possible states), once properly spelled out (absolute continuity), justifies common physical intuition (we assume a probability density exists).
+
+\textbf{Choquet theory.} Classical statistical mechanics and probability start by assuming the existence of a state space $X$ and defining ensembles on top of it.
+This cannot be generalized to quantum mechanics and beyond.
+In our theory of ensemble spaces we start with ensembles, and have to show when and how they can be thought, can be represented, as probability measures.
+If we can assume that $E$ is a compact convex set, Choquet theory answers this question.
+
+First, we have to define what it means to represent an ensemble with a probability measure.
+Let $E$ be a convex subset of a locally convex space $V$ and $X$ the set of extreme points.
+In classical mechanics, $X$ will correspond to the set of Dirac measures, as these cannot be seen as a mixture of other measures.
+In quantum mechanics, $X$ corresponds to the pure states, as these cannot be seen as a mixture of other states.
+In other words, the set of states $X$ is recovered as the set of ensembles with the narrowest spread.
+We say that $\ens \in E$ is represented by a probability measure $p : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_E f dp$ for every continuous linear functional $f$ on $V$.
+Moreover, $p$ is supported by the extreme points if $p(X) = 1$.
+Physically, we start with an ensemble $\ens$ taken as a point in our ensemble space, identified by the value of all possible statistical variables $f(\ens)$, and we represent it as a probability measure $p$, a statistical mixture, over other the extreme points, the states, recovering the expectation with an integral.
+
+We should note that, given any probability measure $p$ over the extreme points $X$, we can always find an $\ens \in E$ that is represented by $p$.
+Therefore, we have a map from the set of probability measures to the set of ensembles.
+The first question, then, is whether we can go back.
+Choquet theory tells us that, in the compact case, we can:
+\begin{thrm}[Choquet 1]
+	Let $E$ be a metrizable compact convex subset of a locally convex space $V$, and let $\ens \in E$.
+Then there exists a probability measure $p_{\ens}$ on $X$ which represents $\ens$ and is supported by the extreme points $X$.
+\end{thrm}
+Note that an ensemble space is a metrizable convex subset of a locally convex TVS. Therefore, if $E$ is compact, the Choquet theorem applies.
+
+The above theorem tells us that the map between probability measures and ensembles is surjective. The second question, then, when is it bijective? Again, Choquet theory has a clear answer.
+\begin{thrm}[Choquet 2]
+	Let $E$ be a closed convex metrizable subset of a locally convex space $V$.
+Then $E$ is a Choquet simplex if and only if for each $\ens \in E$ there exists a unique measure $p_{\ens}$ which represents $\ens$ and is supported by the extreme points $X$.
+\end{thrm}
+The Choquet simplex is a generalization of the finite dimensional simplex, whose mathematical characterization is rather technical.
+The key point is that $E$ is a Choquet simplex if and only if the ensemble space does not allow multiple decomposition in terms of pure states.
+This confirms the physical intuition from classical and quantum statistical mechanics.
+
+For example, the ensemble space for a two state quantum system, the Bloch ball, is a compact convex subset.
+Any internal point can be seen as the convex combination of points on the surface, the pure states.
+The center, which represents the maximally mixed state, can be thought as an equal mixture of any two opposite points, or a uniform distribution over the whole surface.
+The fact that any ensemble can be seen as a mixture of a set of pure states, then, is reflected by the first part of the Choquet theorem, and it is valid for all physical theories.
+Uniqueness of the representation fails, however, precisely because the ensemble space is not a simplex: it is not a classical space.
+
+The core insight from Chouqet theory, then, is that both in the classical and quantum case we can represent ensembles as a probability distribution over a set of states.
+However, only in the classical case is this representation unique.
+In the quantum case, mixed states can be decomposed into pure states in multiple ways, and it is this multiple decomposition that makes the ensemble space not a simplex, and therefore the space not classical.
+Unique representation of ensembles is a desirable property, and standard probability measures are not going to provide it in general.
+
+\section{Quasiprobability}
+
+%Goal of this section: quasi-probability representations are allowed by the embedding in a topological vector space.
+Choice of representations are guided by what observables are interesting for a specific problem.
+
+%Examples: Wigner, Husimi, Glauber P, SIC-POVM (space not made of spectra of operators), Kirkwood-Dirac (two-time correlation function - observables not necessarily taken at the same time )
+
+%Things to say: some people think about it as weakining axioms (cite papers on negative probability).
+It's more than that.
+Points describe events that are not necessarily mutually exclusive.
+TODO: example of two Wigner function/Husimi Q would be better that have no overlap but they are not orthogonal and/or two functions that have overlap but are orthogonal.
+
+%Thing to say: all these representations are useful because they preserve the linear structure, meaning that convex combinations of ensembles are convex combinations of quasi-probability; expectation values are linear.
+Check, but it should be true.
+
+% 1 describe quasiprobability approaches; go through examples; point out that the quasi-probability can be negative; that one can recover expectations;
+
+\textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions, of which the Wigner function is probably the most widely known.
+The Wigner transform $W : \Ens \to L^1(\mathbb{R}^2, \mathbb{R}) $, defined as \cite[eq.~11.8-3a]{mandel1995}
+\begin{equation}
+	W(\rho) = W_{\rho} (q, p) = 2 \int_{-\infty}^{\infty} \langle x + y | \rho | x - y \rangle e^{-2 i p y / \hbar} \; \mathrm dy,
+\end{equation}
+takes a mixed state $\rho$ and returns a real function $W_{\rho}$ of phase space that, like a probability density integrates to one, but, unlike a probability density, can be negative.
+Recall that an affine combination is a linear combination where all coefficients sum to one but are not necessarily non-negative.
+Therefore, in general, quasi-probability distributions can be understood as affine combinations.
+Remarkably, expectation values of polynomials of position and momentum in symmetric ordering coincide with expectations over the Wigner function.
+For example
+\begin{equation}
+	\tr \left( \frac{1}{2}[XP+PX] \rho \right) = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} xp \; W_{\rho} (q, p) dq dp \; .
+\end{equation}
+Moreover, the transform is an affine map: $W(p \rho_1 + \bar{p} \rho_2 ) = p W(\rho_1) + \bar{p} W(\rho_2)$.
+
+In some problems, especially in quantum optics, normal ordering and anti-normal ordering are more useful.
+The Glauber P representation and the Husimi Q representation work like the Wigner function except that the expectations of polynomials recover the respective order.
+In other circumstances, quasi-probability distributions over observables different from position and momentum are more useful.
+For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times \cite{lostaglio2023}.
+Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
+For example, the SIC-POVM represents quantum states as affine combinations of pure states $P_i$ that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
+Each ensemble $\rho$ can then be represented as \cite[eqs.
+28-29]{fuchs2004}
+\begin{equation}
+	\rho = \sum_{n=1}^{d^2} \lambda_i P_i = \sum_{n=1}^{d^2} \left[ (d^2 + 1) p_i - \frac{1}{d} \right] P_i \; ,
+\end{equation}
+where $p_i = \tr \left(P_i \rho\right)$.
+
+It should be clear that there are many ways to represent quantum states via quasi-probability distributions, and that the choice of representation is a matter of convenience for the problem at hand.
+All representations, though, have the following features in common.
+First, we have a map $W : \Ens \to Q(Z,\mathbb{R})$ that given each mixed state $\rho \in \Ens$ returns the corresponding quasi-probability function $w(z) \in Q(Z,\mathbb{R})$, where $Q(Z,\mathbb{R}) \subset L^1(Z,\mathbb{R})$ is the function space used by the specific representation.
+The space $Z$, in fact, can change from representation to representation: it is $\mathbb{R}^{2n}$ for a Wigner function, $\mathbb{C}^n$ for the Husimi Q and $[1, d^2] \subset \mathbb{N}$ for the SIC-POVM.
+All representations preserves convex combinations: $W(\sum_i p_i \rho_i) = \sum_i p_i W(\rho_i)$.
+That is, a statistical mixture of quasi-probability distributions represents the statistical mixture of the corresponding states.
+Since the map is linear, and expectation values are linear, one can express expectation values as integrals or sums over the quasi-probability $w(z)$.
+
+\begin{figure}[H]
+	\centering
+
+	% perspective
+	% from scipy.spatial.transform import Rotation as R
+	% R.from_euler("yx", [50, 10], degrees=True).as_euler("xyz", degrees=True).round(0)[::-1]
+	\newcommand{\yaw}{12}
+	\newcommand{\pitch}{49}
+	\newcommand{\roll}{15}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	\newcommand{\radius}{1}
+
+	\begin{subfigure}[c]{0.23\textwidth}
+		\centering
+		\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
+			% Bloch sphere (outline drawn in canvas plane)
+			\draw (0, 0, 0) circle (\radius cm);
+
+			% meridians and equator
+			\begin{scope}[canvas is yz plane at x=0]
+				\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+				\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is zx plane at y=0]
+				\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+				\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is xy plane at z=0]
+				\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+				\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+			\end{scope}
+
+			% tetrahedron
+			\coordinate (A) at (0, 1, 0);
+			\coordinate (B) at (0, -1/3, {sqrt(8/9)});
+			\coordinate (C) at ({sqrt(2/3)}, -1/3, {-sqrt(2/9)});
+			\coordinate (D) at ({-sqrt(2/3)}, -1/3, {-sqrt(2/9)});
+			\begin{scope}[every node/.style={font=\tiny}]
+				\draw (A) -- (B) -- (C) -- cycle;
+				\draw (A) -- (B) -- (D) -- cycle;
+				\draw (A) -- (C) -- (D) -- cycle;
+				\draw (B) -- (C) -- (D) -- cycle;
+%				\shadedraw (A) circle (0.03 cm) node[above] {$| 0 \rangle$};
+%				\shadedraw (B) circle (0.03 cm) node[below] {$\frac{1}{\sqrt{3}} | 0 \rangle + \sqrt{\frac{2}{3}} e^{i\frac{2\pi}{3}} | 1 \rangle$};
+%				\shadedraw (C) circle (0.03 cm) node[above] {$\frac{1}{\sqrt{3}} | 0 \rangle + \sqrt{\frac{2}{3}} e^{i\frac{4\pi}{3}} | 1 \rangle$};
+%				\shadedraw (D) circle (0.03 cm) node[below] {$\frac{1}{\sqrt{3}} | 0 \rangle + \sqrt{\frac{2}{3}} | 1 \rangle$};
+			\end{scope}
+		\end{tikzpicture}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}[c]{0.23\textwidth}
+		\centering
+		\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=0.65]
+			% Bloch sphere (outline drawn in canvas plane)
+			\draw (0, 0, 0) circle (\radius cm);
+
+			% meridians and equator
+			\begin{scope}[canvas is yz plane at x=0]
+				\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+				\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is zx plane at y=0]
+				\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+				\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is xy plane at z=0]
+				\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+				\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+			\end{scope}
+
+			% inscribed tetrahedron
+			\begin{scope}[scale=3, every node/.style={font=\small}]
+				\coordinate (A) at (0, 1, 0);
+				\coordinate (B) at (0, -1/3, {sqrt(8/9)});
+				\coordinate (C) at ({sqrt(2/3)}, -1/3, {-sqrt(2/9)});
+				\coordinate (D) at ({-sqrt(2/3)}, -1/3, {-sqrt(2/9)});
+				\draw (A) -- (B) -- (C) -- cycle;
+				\draw (A) -- (B) -- (D) -- cycle;
+				\draw (A) -- (C) -- (D) -- cycle;
+				\draw (B) -- (C) -- (D) -- cycle;
+			\end{scope}
+		\end{tikzpicture}
+		\caption{}
+	\end{subfigure}
+	\caption{(a) Example of a SIC-POVM on a Bloch ball.
+The four extreme points of the tetrahedron represent the states $P_i$.
+States with positive $\lambda_i$ lie within the tetrahedron, while all other states require negative coefficients.
+(b) The Bloch ball can alternatively be inscribed within a larger tetrahedron to guarantee positive coefficients, but the extreme points are no longer physical states.}
+\end{figure}
+
+% 2 Note that it's not about negative probability; disjoint not mutually exclusive
+
+\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasi-probability \cite{adhikary2020}.
+Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
+This is a mischaracterization.
+For example, the Husimi Q representation provides us with a non-negative quasi-probability function.
+That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.\footnote{Since the Husimi Q is defined as $Q_{\rho} ( \alpha ) = \frac{1}{\pi} \langle \alpha | \rho | \alpha \rangle$, it is proportional to the probability of measuring a coherent (i.e.
+Gaussian state) $| \alpha \rangle$, which is non-negative.
+} Similarly, a SIC-POVM characterizes the Bloch ball by four pure states that form a tetrahedron inscribed in the ball.
+Since we can also inscribe the Bloch ball inside a tetrahedron, we can represent all states of a two-state quantum system by a convex combination of the vertices, leading to a non-negative quasi-probability distribution over four elements.
+
+Conversely, we can use signed measures to represent classical distributions.
+Suppose we have a two state classical discrete space, which means the space of all possible ensembles is spanned by $[p, (1-p)]$.
+Alternatively, we can write it as $p[1,0] + (1-p)[0,1]$ which highlights that we are writing each possible probability distribution in terms of the two possible cases.
+But we could also express each probability distribution in terms of other two distributions, as in $\left(2p - \frac{1}{2}\right)\left[\frac{3}{4}, \frac{1}{4}\right] + \left(\frac{3}{2} - 2p\right)\left[\frac{1}{4}, \frac{3}{4}\right]$.
+The coefficients now span between $\left[-\frac{1}{2}, \frac{3}{2}\right]$, can be negative, but still sum to one.
+That is, in terms of this new affine basis, we need to use a signed probability.
+
+Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measures, the difference between classical probability and quantum quasi-probability does not lie in the use of negative probability.
+Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events, and all mutually exclusive events are represented by disjoint regions.
+Mutually exclusive events can be prepared and observed independently from the others (i.e.
+they are orthogonal in the sense that they maximize entropy increase during mixture); we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasi-probability in quantum mechanics.
+
+For example, given a pure state $|\psi\>$, the Husimi Q function representing it is given by $Q_\psi(\alpha) = \frac{1}{\pi} |\langle \alpha | \psi \rangle|^2$.
+The inner product is given by \cite[eq.
+11.7-1]{mandel1995}
+\begin{equation}
+	\langle \alpha | \psi \rangle = e^{-|\alpha|^2/2} F_\psi(\alpha^\ast)
+\end{equation}
+where $F_\psi$ is an entire analytic function, which cannot vanish on an open set without vanishing everywhere, therefore the support of all Husimi Q functions is the whole space $Z$.
+This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support; moreover disjoint subsets of $Z$ cannot be thought of as mutually exclusive events because we are not going to be able to find two distributions with those subsets as their respective supports.
+For the discrete case, a SIC-POVM representation picks $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot all be orthogonal to each other.
+For a Bloch sphere, the four corners of the tetrahedron are not all opposite to each other.
+Therefore, disjoint sets of $P_i$ do not correspond to mutually exclusive events.
+Quasi-probability representations, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
+
+% 3 These representations are possible because of the vector space structure of the ensemble space, and therefore something similar will be possible in ANY physical theory
+
+\textbf{Ensemble space embedding and signed measures.} We saw that negative probability is a red herring, and mutual exclusivity is key.
+But how exactly does mutual exclusivity give us classical probability?
+Moreover, are quasi-probability representations specific to quantum mechanics, or are they more general?
+This is where the perspective from the ensemble space framework allows us to answer these questions directly.
+
+In Choquet theory, we saw that classical spaces, Choquet simplexes, are those in which there is a one-to-one correspondence between probability measures and ensembles, where no multiple decomposition in terms of pure states is allowed.
+We want to show that this is a consequence of assuming mutual exclusivity between the points of $X$.
+Suppose we have an ensemble $\ens$ that can be represented as different measures $p_1$ and $p_2$ over the extreme points.
+This means that there are two disjoint sets $A, B \subseteq X$ for which $p_1(A) < p_2(A)$ and $p_1(B) > p_2(B)$.
+That is, we can move probability from $A$ to $B$ while leaving the ensemble, and therefore the entropy and all statistical variables, unchanged.
+If $A$ and $B$ were two mutually exclusive events, this would not be possible: if we could distinguish between $A$ and $B$, something would have to change.
+The only way for $A$ and $B$ to be mutually exclusive, then, is if different measures represent different ensembles, which means the ensemble space is a simplex.
+This tells us that we can use classical probability exactly when all extreme points are mutually exclusive (i.e.
+orthogonal as defined from the entropy) and the ensemble space is a simplex.
+
+That is, classical probability does not merely correspond to the use of a single non-negative probability measure.
+It corresponds to the ability to, conceptually, allow any possible probability measure to describe a possible distinct scenario.
+This requirement is exactly mutual exclusivity.
+In quantum mechanics, a classical context is a simplex embedded within the ensemble space.
+For the Bloch ball, a classical context is an axis that connects two opposite points.
+The points are orthogonal because the midpoint is the maximally mixed state, which therefore maximizes the entropy increase, and the line segment is a one simplex.
+If we take the position operator and consider all density operators that commute with $x$, this is corresponds to the set of functions $\rho(x)$ that integrate to one, which is the space of probability supported by $X$, which is a simplex.
+
+Now the second question: why does quantum mechanics allow quasi-probability representations?
+Let $Z$ be the space over which we are defining our quasi-probability distributions.
+Note that, in general, $Z$ is not the space of states.
+Wigner functions, for example, are defined over phase space, which does not represent quantum states.
+Let $M^{\pm}(Z)$ be the space of all bounded signed measures over $Z$.
+Since the linear combination of two bounded signed measures is a bounded signed measure, $M^{\pm}(Z)$ is a vector space.
+Now let $M_1^{\pm}(Z)$ be the space of signed probability measures.
+While it is not a subspace of $M(Z)$, it does not include the zero measure, it is an affine subspace therefore, up to the choice of an arbitrary origin, it is a vector space.
+
+Recall that an ensemble space must be a convex subset of a vector space.
+If we are able to represent ensembles as quasi-probability distributions over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such an $Z$ always exists is still open for us.
+A possible conjecture is that $Z$ could be constructed from the statistical quantities $F_i$ that fully identify each ensemble in $\Ens$.} In this case, any element $v\in V$ can be represented uniquely by a signed probability measure $\mu_{v}$, regardless of whether it is an actual ensemble or not.
+Since statistical quantities $F_i$ are linear functional, we will have $F_i(v) = \int_Z F_i(z) d\mu_{v}$.
+In other words, there is nothing special about quantum mechanics that allows quasi-probability representations.
+It is the fact that the ensemble space embeds into a vector space, a general fact for all theories, that allows such representations.
+
+It should be clear that, since an ensemble space is a strict subset of the embedding vector space, not all affine measures will correspond to an ensemble space.
+Whether all probability measures correspond to a valid ensemble, or whether signed probability measures are even used depends on how the mapping is done in each particular case.
+For example, there is no simple characterization of which Wigner functions correspond to valid quantum states and which do not.
+
+% 4 pros and cons - can choose a repretation for the problem at hand; no fundamental representation; boundaries ore not well defined (i.e. which functions represnt valid states)
+
+\textbf{Non-uniqueness of quasi-probability.} Our approach gives us a consistent view of both quasi-probabilities and classical probability.
+The first are affine representations of the ensembles that connect to the vector space structure, while the second are convex representations of the extreme points which rely on the shape of the ensemble space.
+
+Quasi-probability distributions are very useful in practice, particularly because one can construct them to fit particular problems.
+Once one has identified a set of observables that are the most natural ones to define a problem, one may find a representation that is specifically suited for the case.
+While the quasi-probability cannot be interpreted as probability, it still provides a useful tool to describe and understand the problem and give some intuition.
+
+However, this practical convenience is also a drawback for finding a more general approach to probability in spaces beyond classical mechanics.
+Crucially, the space $Z$ over which the quasi-probability measures are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tell us which quasi-probability measures represent valid ensembles and which do not.
+
+As a comparison, in the previous case all probability measures were given over a uniquely defined set of points, the extreme points of the ensemble space, but the probability measure may not be unique.
+In this case, the quasi-probability associated to each ensemble is unique, but the space over which the quasi-probability is defined is not.
+
+\section{Fraction capacity}
+
+\textbf{Probability of outcomes vs fraction of components.} We saw that classical probability fails in quantum mechanics because pure states are not always mutually exclusive, are not orthogonal.
+That is why, in quantum mechanics, probability can be defined only once a context, an orthogonal basis, is chosen.
+Therefore, at a conceptual level, probabilities can only give us a patchwork representation of ensembles.
+Some authors embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.
+[CITATION NEEDED] However, there is a different route.
+
+Note that the linearity in our axioms is in terms of mixing coefficients, not probability of outcomes.
+That is, regardless of whether we can characterize experiment outcomes with probability distributions, preparation procedures will always allow us to create a statistical mixture.
+As we saw in Choquet theory, the difficulty is that there are multiple ways to mix pure states to recover the same ensemble.
+However, the question of what is the biggest mixture coefficient one can use in a specific case is well-defined, and generalizes to the non-compact case as well.
+Therefore, our generalization of probability is not in terms of what an ensemble can lead to (i.e.
+measurements), but in terms of how ensembles can be produced (i.e.
+preparations).
+
+\textbf{Fraction capacity.} Let us concentrate on the following problem: given an ensemble $\ens \in \Ens$ and a (Borel) set of ensembles $A \subseteq \Ens$, what fraction of $\ens$ can be constructed by a mixture of $A$?
+How much of $\ens$ can be explained as coming from preparations corresponding to $A$?
+For example, let $\Ens$ be the space of all probability distributions for a six-sided die.
+Let $\ens_{123456}$ be the uniform distribution over all outcomes.
+Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
+Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$, where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/6$ of $\ens$ can be constructed from $A$, but no more.
+Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
+Note that, given the uniform distribution, $1/6$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
+This gives the basic insight for our definitions.
+
+Given a target ensemble $\ens \in \Ens$ and an arbitrary ensemble $\ens[a] \in \Ens$, we define the \textbf{fraction} of $\ens[a]$ in $\ens$ to be
+\begin{equation}
+	\fraction_{\ens}(\ens[a]) = \sup(\{ p \in [0,1] \, | \, \exists \, \ens_1 \in \Ens \text{ s.t. }  \ens = p \ens[a] + \bar{p} \ens_1 \}).
+\end{equation}
+The fraction is always well-defined because we can always write $\ens = 0 \ens[a] + 1 \ens$, therefore it must be zero or greater.
+This quantity tells us how much of the ensemble $\ens$ can be constructed from $\ens[a]$.
+
+We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a Borel set of ensembles $A \subset \Ens$.
+Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.
+mixtures),\footnote{Since we work with convex sets within a topology, the hull is both closed under the topology and convex combinations.
+That is, $\hull(A)$ is the smallest closed and convex set that contains $A$} we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
+\begin{equation}
+	\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\}).
+\end{equation}
+This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$.
+
+\textbf{Non-additivity of the fraction capacity.} One can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
+\begin{enumerate}
+	\item non-negative and unit bounded - $\frcap_{\ens}(A) \in [0,1]$
+	\item monotone - $A \subseteq B \implies \frcap_{\ens}(A) \leq \frcap_{\ens}(B)$
+	\item sub-additive - $\frcap_{\ens}(A \cup B) \leq \frcap_{\ens}(A) + \frcap_{\ens}(B)$
+	\item continuous from below and above - $\frcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \frcap_{\ens}(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
+\end{enumerate}
+
+The first property readily comes from the range of the fraction.
+For the second, note that $\hull$ and $\sup$ are both monotone.
+For sub-additivity, note that the fraction of two components can at most sum during mixing.
+For the last, increasing and decreasing sequences strictly add or remove possible mixtures, therefore they will lead to increasing or decreasing sequences of real numbers bounded between zero and one.
+The limit of these sequences must agree with the fraction capacity of the limit.
+
+Note that, if not for the lack of additivity over disjoint sets, the fraction capacity would be a probability measure.
+It is therefore a non-additive probability measure.
+The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
+fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\footnote{ \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} }
+
+\textbf{Compare to Choquet theory.} In case $E$ is compact, we can restrict the fraction capacity to the set $X$ of extreme points.
+Given a set $A \subset X$, of extreme points, we can find the set $N \subset E$ of all ensembles that can be represented with a probability measure with support $A$.
+This forms a convex set.
+Now, the ability to express an ensemble $\ens$ as the convex combination of $N$ is precisely the ability to find a measure $p$ that represents $\ens$ such that $p(A) \neq 0$.
+That is, $\ens = \lambda \ens[a] + \bar{\lambda} \ens[b]$ with $\ens[a] \in M$ if and only if there is a probability measure $p$ that represents $\ens$ such that $p(A) \geq \lambda$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{p(A) \, | \, p \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all possible measures.
+
+As we saw, the classical case corresponds to a simplex in Choquet theory.
+This means that each element of the ensemble space is represented by a unique probability measure over the extreme points.
+In that case, the fraction capacity restricted to the extreme points will also be additive.
+In the non-classical case, the closure of the ensemble space will not be a simplex, and there will be multiple representations, which will lead to a non-additive measure.
+Suppose, in fact, that $p_1$ and $p_2$ are two distinct measures that represent the same ensemble $\ens$.
+Then there will be a set $A$ such that $p_1(A) \neq p_2(A)$.
+Suppose, without loss of generality, that $p_1(A) > p_2(A)$ We will also have $p_1(A^{\complement}) = 1- p_1(A) < 1 - p_2(A) = p_2(A^{\complement})$.
+Since the fraction capacity for a set cannot be less than the measure on that set, we have $\frcap_{\ens}(A \cup A^{\complement})=\frcap_{\ens}(X) = 1 = p_1(A) + p_1(A^{\complement}) < (A) + p_2(A^{\complement}) \leq \frcap_{\ens}(A) + \frcap_{\ens}(A^{\complement})$.
+That is, the fraction capacity is additive over disjoint sets of extreme points in and only in the classical case.
+In other words, the fraction capacity reduces to a classical probability distribution over pure states in the classical case.
+
+\textbf{Uniqueness of the fraction capacity.} From a conceptual standpoint, the fraction capacity has a major advantage: it has a clear physical meaning and it is unique.
+Like in Choquet theory and unlike for quasi-probability, the domain over which the measure is defined is fully determined by the system, by the ensemble space.
+On the other hand, unlike Choquet theory and like quasi-probability, once the space is fixed, there is one unique fraction capacity for each ensemble.
+Moreover, the setup is more general than Choquet theory (i.e.
+the ensemble space need not be compact) and stems from necessary physical requirements.
+An in depth study on fraction capacities, then, is guaranteed to be applicable to all physical theories, including those yet to be formulated.
+
+One important feature, for example, would be to characterize what fraction capacities are possible for each ensemble space.
+In the classical case, for example, only additive fraction capacities are possible; in the quantum case, only some type of non-additivity will be possible.
+The shape of the space is effectively encoded in the space of possible fraction capacities, maybe in a way that is mathematically and physically insightful.
+
+\section{State capacity}
+
+\textbf{Counting states.} Since the fraction capacity can be seen as the non-additive generalization of a probability measure, we now want to introduce a similar generalization for the notion of count of states.
+Since entropy characterizes the variability of the preparations of the ensemble, greater variability means the ensemble is spread over more cases.
+Therefore the count of distinguishable states must be a monotonic function of the entropy.
+Given that in statistical mechanics the entropy is the logarithm of the count of states, we expect the count of states to be the exponential of the entropy.
+Therefore one can argue that the proper way to measure the spread of an ensemble is through the exponential of the entropy, also known as entropic volume.\footnote{It has already been shown in \cite{hall1999} that the exponential of the entropy is the only function that can be used to define a volume.} However, the count of states, like the probability measure, is typically defined on the sample space $\Omega$, which can only work in the classical case.
+The entropy volume is a function of the ensemble space, so it is not a measure.
+Let's see how everything fits together.
+
+To build an intuition, consider the classical case.
+For every region $U$ of phase space, we have a set of possible probability distributions that have support $U$.
+This is a convex set since every mixture of distributions with support $U$ has support $U$.
+Note that the uniform distribution $\rho_U$, if it exists, is the one with the greatest entropy, which will correspond to $S(\rho_U) = \log \mu(U)$.
+If $U$ is not compact, then there is no upper limit to the entropy and $\mu(U) = +\infty$.
+Now, let us transfer this intuition on the ensemble space $\ens$.
+If we take a set $A$, we can ask: what is the highest entropy reachable by a mixture of elements of $A$?
+If $A$ happens to be the set of ensembles with support $U$, then the supremum of the entropy is $\log \mu(U)$, which also works in the unbounded case.
+
+The above intuition started with the notion of the state space $X$ and the count of states $\mu$.
+In the general case, we will start with the ensemble space $\Ens$ and the entropy $S$.
+For ensembles, we can define the entropic volume\footnote{CITE Hall} to be the exponential of the entropy $2^{S(\ens)}$.
+For the entropic volume to quantify the spread of the ensemble, we must have that the spread can at most sum.
+Remarkably, the bounds on the entropy are such that the exponential of the entropy is subadditive.
+That is, $2^{S(p\ens[a]+\bar{p}\ens[b])} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])}$ for any ensemble space.\footnote{It is an open question whether the exponential is the only monotonic function of the entropy that is subadditive.} For the equality to hold, the ensemble must be orthogonal, meaning that two states are distinguishable if and only if they are mutually exclusive.
+Therefore, this is indeed a general way to define the spread of ensembles and the count of distinguishable states in a way that is consistent with the previous discussions.
+
+\textbf{State capacity.} Given a Borel set of ensembles $A \subseteq \Ens$, we define the \textbf{state capacity} of $A$ to be
+\begin{equation}
+\capacity(A) = \sup(2^{S(\hull(A))}\cup\{0\}).
+\end{equation}
+This returns the upper bound to the exponential of entropy achievable with a mixture of elements of $A$.
+
+One can then show that the fraction capacity $\capacity : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
+\begin{enumerate}
+	\item non-negative - $\capacity(A)\geq 0$
+	\item monotone - $A \subseteq B \implies \capacity(A) \leq \capacity(B)$
+	\item sub-additive - $\capacity(A \cup B) \leq \capacity(A) + \capacity(B)$
+	\item additive over orthogonal sets - $\capacity(A \cup B) = \capacity(A) + \capacity(B)$ if $A \ortho B$
+	\item continuous from below and above - $\capacity(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \capacity(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
+\end{enumerate}
+
+The first property readily comes from the range of the exponential.
+For the second, note that $\hull$, exponential and $\sup$ are all monotone.
+For sub-additivity, the exponential of the entropy itself is sub-additive, and it is additive over orthogonal ensembles.
+For the last, since the entropy itself  is continuous, this property carries over to the set function.
+
+Again we note that the non-additive measure we recover is very well-behaved, and connects directly with important quantities.
+In the compact case, state capacity can be extended to the extreme points.
+In the classical case, disjoint sets of extreme points will be orthogonal, meaning that the state capacity there restricted is an additive measure.
+In the discrete case, every point is associated with zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
+In the continuous case, gives a subset of phase space $U$ and $A_U$ the set of distributions with support within $U$, the maximum entropy is reached by the uniform distribution $\rho_U \in A_U$, which means we recover $\log_2 \capacity(A_U) = S(\rho_U)$, the fundamental postulate of statistical mechanics.
+Similarly, in quantum mechanics, given an orthogonal subspace $A$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(A) = 2^{\dim(A)}$.
+
+The state capacity, then, recovers the number of distinguishable states.
+In classical mechanics, all states are perfectly distinguishable as they represent mutually exclusive cases, the measure is additive.
+In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states.
+We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutual exclusiveness, is responsible for the failure of the additivity of the measure.
+Note that, since the fraction capacity depends only on the ensemble space, and not also on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
+
+\section{Non-additive generalization of probability theory}
+
+\textbf{Requirements for non-additive probability theory.} In a previous section we reviewed what types of objects we have in probability theory for classical ensembles: a sample space $X$, a probability measure $p$, random variables $F : X \to \mathbb{R}$, expectations as integrals of the probability measure $E[F] = \int_X F(x) dp$, a state counting measure $\mu$, the probability density as the Radon-Nikodym derivative $\frac{dp}{d\mu}$ and the entropy as $\int_X \rho \log \rho d\mu$.
+Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of the probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the states counting measure $\mu$, is there a generalization of all other elements?
+That is, can we find an integral and derivative operation over non-additive measures that generalizes to all physical theories?
+
+The first difficulty, as we already pointed out, is that the ensemble space $\Ens$ may not have extreme points.
+What is the correct procedure to ``close'' the ensemble space and recover the space of extreme points is, for us, an open question.
+For now, we will assume $E$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
+Therefore we can talk about the restriction on $X$ of the state capacity and the fraction capacity.
+
+Recall that a statistical variable is a linear function $F : E \to \mathbb{R}$ defined over the ensemble space.
+Given that $F$ is a linear, its restriction on $X$ will determine the values for all other points of $E$.
+Since the expectation is recovered through the integration of the variable over the probability density, we would like a non-additive integral $(E)\int_X$ such that
+\begin{equation}
+	E[F] = F(\ens) = (E)\int_X F(x) d\frcap_{\ens}.
+\end{equation}
+which reduces to the Riemann integral in the classical (additive) case.
+The integral will depend only on the values of $F$ and $\frcap_{\ens}$ on $X$, but since these uniquely determine both $F$ and $\ens$, the integral can be understood as a map from $F$ and $\ens$ to $F(\ens)$, which is, in principle, well-defined.
+Therefore, in principle, this integral should exist.
+
+Additionally, we would like to be able to define an analogue of the probability density.
+We are therefore looking for a non-additive derivative $\rho = \frac{d \frcap_{\ens}}{d \capacity}$ that reduces to the Radon-Nikodym derivative in the classical (additive) case such that
+\begin{equation}
+	F(\ens) = (E)\int_X F(x) \rho(x) d\capacity
+\end{equation}
+under the previous integral. It is less clear whether this derivative is always well defined. Ideally, we would also expect
+\begin{equation}
+	S(\ens) = (E)\int_X \rho(x) \log \rho(x) d\capacity.
+\end{equation}
+This last requirement may not be possible, or it may be used as a guide in case the previous derivative is not uniquely defined.
+
+\textbf{Ruling out Sugeno and Choquet integrals.} Given that extensive literature already exists for non-additive measure theory (cite), as it is used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence, we originally thought it would be just a matter of clarifying our definitions and connecting them to the existing literature.
+However, things turned out to be more complicated.
+Depending on the use cases, different types of integrals exist.
+In fact, procedures exist\cite{nonadditive2014,pap2013} to construct infinitely many types of non-additive integrals.\footnote{Briefly, one can define a pseudo-addition and a pseudo-multiplication as real functions of two real arguments and construct an integral based on those definitions.} At this point, we can rule out the two most common non-additive integrals: the Sugeno integral and the Choquet integral.
+
+The Sugeno integral does not recover the Lebesgue integral in the case the capacity is additive (i.e.
+classical), and is therefore quickly ruled out.
+The Choquet integral does, so it will work in the classical case.
+However, it does not work in the quantum case.
+We can show this by computing the Choquet integral for a two state quantum system and see that it does not recover the expectation of the statistical variables in the corresponding ensemble space.
+
+\begin{figure}[H]
+	\centering
+
+	% perspective
+	\newcommand{\yaw}{11}
+	\newcommand{\pitch}{28}
+	\newcommand{\roll}{23}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	% dimensions
+	\newcommand{\s}{-0.5}
 	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
@@
 	\newcommand{\margin}{1.3}
+
+	\definecolor{colora}{HTML}{1EBCB4}
+	\definecolor{colorb}{HTML}{E57C04}
+	\definecolor{colorc}{HTML}{7A5CE3}
 
@@
 
-		% cut
+		% cut 1
+		\begin{scope}[canvas is zx plane at y=\radius-\s]
+			\draw (-0.9*\margin, -1.1*\margin) node[below left, circle, draw, thick, color=colora] {1};
+			\filldraw[color=colora, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+		\end{scope}
+
+		% cut 2
+		\begin{scope}[canvas is zx plane at y=-\s]
+			\draw (-0.9*\margin, -1.1*\margin) node[below left, circle, draw, thick, color=colorb] {2};
+			\filldraw[color=colorb, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[color=colorb, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[color=colorb, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+		\end{scope}
+
+		% cut 3
 		\begin{scope}[canvas is zx plane at y=\s]
-			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
-			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw (-0.9*\margin, -1.1*\margin) node[below left, circle, draw, thick, color=colorc] {3};
+			\filldraw[color=colorc, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[color=colorc, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[color=colorc, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
 		\end{scope}
@@
 		% empty case
-		\draw (3.75*\xx+\xy, 3.75*\yx+\yy, 3.75*\zx+\zy) node {$s > b$};
+		\draw[color=colora] (3.75*\xx+\xy, 3.75*\yx+\yy, 3.75*\zx+\zy) node {$s > b$};
 
@@
 		\begin{scope}[shift={(4.5*\xx-0.2*\xy, 4.5*\yx-0.2*\yy, 4.5*\zx-0.2*\zy)}, scale=0.5]
-			\draw (-3.5, +1.3, 0) node {$\frac{1}{2} ( a + b ) < s < b$};
+			\draw[color=colorb] (-3.5, +1.3, 0) node {$\frac{1}{2} ( a + b ) < s < b$};
 
@@
 			\begin{scope}[canvas is zx plane at y=-\s]
-				\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-				\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[color=colorb, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[color=colorb, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
 			\end{scope}
@@
 		\begin{scope}[shift={(3*\xx-\xy, 3*\yx-\yy, 3*\zx-\zy)}, scale=0.5]
-			\draw (3.5, -0.5, 0) node {$s < \frac{1}{2} ( a + b )$};
+			\draw[color=colorc] (3.5, -0.5, 0) node {$s < \frac{1}{2} ( a + b )$};
 
@@
 			\begin{scope}[canvas is zx plane at y=\s]
-				\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-				\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[color=colorc, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[color=colorc, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
 			\end{scope}
@@
 On the left, the Bloch ball with a plane at constant $s$.
-On the right, the top case is when the value is greater than the maximum (empty set); the middle case is when the value is in between the target point $\ens[o]$ and the maximum; the third case is when the value is below the target point.}
+On the right, the top case is when the value is greater than the maximum (empty set); the middle case is when the value is between the target point $\ens[o]$ and the maximum; the third case is when the value is below the target point.}
 \end{figure}
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{ChoquetCalculation2.jpg}
-	\caption{...}
+
+	% perspective
+	\newcommand{\yaw}{11}
+	\newcommand{\pitch}{28}
+	\newcommand{\roll}{23}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	% dimensions
+	\newcommand{\s}{0.5}
+	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
+	\pgfmathsetmacro{\stheta}{asin(\s)}
+	\pgfmathsetmacro{\minstheta}{asin(-\s)}
+
+	\newcommand{\radius}{1}
+	\newcommand{\margin}{1.3}
+
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
+		% Bloch sphere (outline drawn in canvas plane)
+		\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, radius=1 cm] coordinate (A);
+%		\draw[dash dot, red] (A) arc[start angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, end angle=-90, radius=1 cm];
+		\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, radius=1 cm] coordinate (B);
+%		\draw[dash dot, green] (B) arc[start angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, end angle=270, radius=1 cm];
+		\draw[dashed] (A) arc[start angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)-360}, radius=1 cm];
+
+		% special points
+		\shadedraw (0, 0, 0) circle (0.03 cm) node[below right] {$\mathsf{o}$};
+		\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+		\shadedraw (0, -\radius, 0) circle (0.03cm) node[below] {$\mathsf{a}$};
+		\shadedraw (0, \s, 0) circle (0.03cm) node[left] {$\mathsf{e}$};
+
+		% convex combination
+		\draw (0, -\radius, 0) -- (0, \s, 0);
+		\draw (2.25*\xx, 2.25*\yx, 2.25*\zx) node {$\ens[o] = \frac{\overline{\ens[o]\ens[a]}}{\overline{\ens[e]\ens[a]}} \ens[e] + \frac{\overline{\ens[o]\ens[e]}}{\overline{\ens[e]\ens[a]}} \ens[a]$};
+
+		% s axis
+		\begin{scope}[shift={(-1.5*\xx, -1.5*\yx, -1.5*\zx)}]
+			\draw (0, -1, 0) -- (0, 0, 0);
+			\draw[->] (0, 0, 0) -- (0, 1, 0) node[below left] {$s$};
+		\end{scope}
+
+		% cut
+		\begin{scope}[canvas is zx plane at y=\s]
+%			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+		\end{scope}
+
+		% meridians
+		\begin{scope}[canvas is yz plane at x=0]
+			% foreground above s
+			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=90-\stheta, radius=1 cm] coordinate (C);
+			% foreground below s
+			\draw[dashed] (C) arc[start angle=90-\stheta, end angle={atan2(-\ixz, -\ixy)}, radius=1 cm] coordinate (D);
+			% background below s
+			\draw[loosely dotted] (D) arc[start angle={atan2(-\ixz, -\ixy)}, end angle=\stheta+270, radius=1 cm];
+			% background above s
+			\draw[densely dotted] (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=\stheta-90, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is xy plane at z=0]
+			% foreground above s
+			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=180-\stheta, radius=1 cm] coordinate (E);
+			% foreground below s
+			\draw[dashed] (E) arc[start angle=180-\stheta, end angle={atan2(\izy, \izx)+360}, radius=1 cm] coordinate (F);
+			% background below s
+			\draw[loosely dotted] (F) arc[start angle={atan2(\izy, \izx)}, end angle=\stheta, radius=1 cm];
+			% background above s
+			\draw[densely dotted] (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=\stheta, radius=1 cm];
+		\end{scope}
+	\end{tikzpicture}
+	\caption{Geometric representation of the element $\ens[e]$ in $A(s)$ with the largest fraction of $\ens[o]$.}
 \end{figure}
@@
 In this case, the hull of $A(s)$ is the top spherical cap.
-We need to consider all possible convex combinations $\ens[o] = p \ens + \bar{p} \ens[c]$ where $\ens$ is in the spherical cap and $\ens[c]$ is not, and find the one with highest $p$.
+We need to consider all possible convex combinations $\ens[o] = p \ens + \bar{p} \ens[c]$ where $\ens$ is in the spherical cap and $\ens[c]$ is not, and find the one with the highest $p$.
 Since, for combinations along the same direction, $p$ increases as $\ens$ gets closer to $\ens[o]$ and as $\ens[c]$ gets further, $\ens[c]$ will be an extreme point on the lower hemisphere while $\ens$ will be on the disc at constant $s$, the lower side of the cap.
@@
 
-To recap, we have (TODO: add a graph)
+To recap, we have
 \begin{equation}
@@
 \end{equation}
+\begin{figure}[H]
+	\centering
+	\newcommand{\ticklength}{0.03}
+	\begin{tikzpicture}[scale=3, xticklabel/.style={anchor=base, inner sep=1pt, yshift=-2.5ex}]
+		% axes
+		\draw[->] (0, 0) -- (1.7, 0);
+		\draw (0, 0) node[left] {$0$} -- +(-\ticklength, 0) (0, 0.25) node[left] {$1/2$} -- +(-\ticklength, 0) (0, 0.5) node[left] {$1$} -- +(-\ticklength, 0);
+		\draw[->] (0, 0) -- (0, 0.6);
+		\draw (0, 0) node[xticklabel] {$0$} -- +(0, -\ticklength) (0.5, 0) node[xticklabel] {$a$} -- +(0, -\ticklength) (1, 0) node[xticklabel] {$1/2(a+b)$} -- +(0, -\ticklength) (1.5, 0) node[xticklabel] {$b$} -- +(0, -\ticklength);
+		% cases s <= 1/2(a+b) and 1/2(a+b) < s <= b
+		\draw[thick, blue, domain=1:1.5, samples=200] (0, 0.5) --plot (\x, {0.25/(\x-0.5)});
+		% mark discontinuity
+		\filldraw (1.5, 0.25) circle (0.02cm);
+		\draw (1.5, 0) circle (0.02cm);
+		% case s > b
+		\draw[thick, blue] (1.5, 0) -- (1.7, 0);
+	\end{tikzpicture}
+	\caption{The integrand $\frcap_{\ens[o]}(A(s))$ as a function of $s$.}
+\end{figure}
 We can now calculate the integral

### 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: two typos
Two more obvious typos I missed.

--- 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-TODO: figure with classical triangle and quantum Bloch Ball
+\begin{figure}[H]
+	\centering
+
+	% perspective
+	\newcommand{\yaw}{11}
+	\newcommand{\pitch}{28}
+	\newcommand{\roll}{23}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	% dimensions
+	\newcommand{\radius}{1}
+
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
+		% Triangle
+		\draw (canvas cs: x=-3cm, y=-0.25cm) +(canvas polar cs: angle=90, radius=1cm) -- +(canvas polar cs: angle=210, radius=1cm) -- +(canvas polar cs: angle=330, radius=1cm) -- cycle;
+
+		% Bloch sphere (outline drawn in canvas plane)
+		\draw (0, 0, 0) circle (\radius cm);
+
+		% meridians and equator
+		\begin{scope}[canvas is yz plane at x=0]
+			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+			\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is zx plane at y=0]
+			\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+			\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is xy plane at z=0]
+			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+		\end{scope}
+	\end{tikzpicture}
+	\caption{Examples of a classical discrete ensemble space with three pure states (triangle) and of a two dimensional quantum ensemble space (ball).}
+\end{figure}
 
@@
 Therefore, at a conceptual level, probabilities can only give us a patchwork representation of ensembles.
-Some author embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.
+Some authors embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.
 [CITATION NEEDED] However, there is a different route.
@@
 In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states.
-We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutually exclusiveness, is responsible for the failure of the additivity of the measure.
+We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutual exclusiveness, is responsible for the failure of the additivity of the measure.
 Note that, since the fraction capacity depends only on the ensemble space, and not also on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{ChoquetCalculation2.jpg}
-	\caption{...}
+
+	% perspective
+	\newcommand{\yaw}{11}
+	\newcommand{\pitch}{28}
+	\newcommand{\roll}{23}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	% dimensions
+	\newcommand{\s}{0.5}
+	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
+	\pgfmathsetmacro{\stheta}{asin(\s)}
+	\pgfmathsetmacro{\minstheta}{asin(-\s)}
+
+	\newcommand{\radius}{1}
+	\newcommand{\margin}{1.3}
+
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
+		% Bloch sphere (outline drawn in canvas plane)
+		\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, radius=1 cm] coordinate (A);
+%		\draw[dash dot, red] (A) arc[start angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, end angle=-90, radius=1 cm];
+		\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, radius=1 cm] coordinate (B);
+%		\draw[dash dot, green] (B) arc[start angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, end angle=270, radius=1 cm];
+		\draw[dashed] (A) arc[start angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)-360}, radius=1 cm];
+
+		% special points
+		\shadedraw (0, 0, 0) circle (0.03 cm) node[below right] {$\mathsf{o}$};
+		\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+		\shadedraw (0, -\radius, 0) circle (0.03cm) node[below] {$\mathsf{a}$};
+		\shadedraw (0, \s, 0) circle (0.03cm) node[left] {$\mathsf{e}$};
+
+		% convex combination
+		\draw (0, -\radius, 0) -- (0, \s, 0);
+		\draw (2.25*\xx, 2.25*\yx, 2.25*\zx) node {$\ens[o] = \frac{\overline{\ens[o]\ens[a]}}{\overline{\ens[e]\ens[a]}} \ens[e] + \frac{\overline{\ens[o]\ens[e]}}{\overline{\ens[e]\ens[a]}} \ens[a]$};
+
+		% s axis
+		\begin{scope}[shift={(-1.5*\xx, -1.5*\yx, -1.5*\zx)}]
+			\draw (0, -1, 0) -- (0, 0, 0);
+			\draw[->] (0, 0, 0) -- (0, 1, 0) node[below left] {$s$};
+		\end{scope}
+
+		% cut
+		\begin{scope}[canvas is zx plane at y=\s]
+%			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+		\end{scope}
+
+		% meridians and equator
+		\begin{scope}[canvas is yz plane at x=0]
+			% foreground above s
+			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=90-\stheta, radius=1 cm] coordinate (C);
+			% foreground below s
+			\draw[dashed] (C) arc[start angle=90-\stheta, end angle={atan2(-\ixz, -\ixy)}, radius=1 cm] coordinate (D);
+			% background below s
+			\draw[loosely dotted] (D) arc[start angle={atan2(-\ixz, -\ixy)}, end angle=\stheta+270, radius=1 cm];
+			% background above s
+			\draw[densely dotted] (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=\stheta-90, radius=1 cm];
+		\end{scope}
+%		\begin{scope}[canvas is xy plane at z=0]
+%			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+%			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+%		\end{scope}
+		\begin{scope}[canvas is xy plane at z=0]
+			% foreground above s
+			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=180-\stheta, radius=1 cm] coordinate (E);
+			% foreground below s
+			\draw[dashed] (E) arc[start angle=180-\stheta, end angle={atan2(\izy, \izx)+360}, radius=1 cm] coordinate (F);
+			% background below s
+			\draw[loosely dotted] (F) arc[start angle={atan2(\izy, \izx)}, end angle=\stheta, radius=1 cm];
+			% background above s
+			\draw[densely dotted] (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=\stheta, radius=1 cm];
+		\end{scope}
+	\end{tikzpicture}
+	\caption{Geometric representation of the element $\ens[e]$ in $A(s)$ with the largest fraction of $\ens[o]$.}
 \end{figure}

### 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: grammar
I incorporated some grammatical suggestions from Grammarly.

@@
 Moreover, we will also assume $E$ to be compact, even though it will not be in general.
-How to properly make these identifications, and how to go beyond the compact case, is outside of the scope of this particular article.\footnote{An early attempt used the notion of orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
+How to properly make these identifications, and how to go beyond the compact case, is outside the scope of this particular article.\footnote{An early attempt used the notion of orthogonality to create the lattice of orthogonal subspaces and use something along the lines of Stone's representation theorem for Boolean algebra.
 This does not work.
@@
 
-It is the ease with which such distinctions between the classical and quantum case can be made that is the main appeal for this structure.
+It is the ease with which such distinctions between the classical and quantum case can be made that is the main appeal of this structure.
 We will see over and over how mutual exclusivity is linked to classical probability and its failure in quantum, and other non-classical, theories.
@@
 the sample space), ensembles are defined as probability measures $p : \Sigma_X \to [0,1]$.
-That is, a probability measure assigns to each element of the Borel algebra $\Sigma_X$\footnote{In our overall project, Borel sets represent statements associated to a test, regardless of termination.}, an event, the probability to find it to be true.
-In the discrete case, like for roll of a die, $X$ is a set endowed with the discrete topology\footnote{Physically, the discrete topology corresponds to the ability to experimentally verify or falsify each case.
+That is, a probability measure assigns to each element of the Borel algebra $\Sigma_X$\footnote{In our overall project, Borel sets represent statements associated with a test, regardless of termination.}, an event, the probability to find it to be true.
+In the discrete case, like for rolling a die, $X$ is a set endowed with the discrete topology\footnote{Physically, the discrete topology corresponds to the ability to experimentally verify or falsify each case.
 That is, we are able to recognize whether a particular die landed or didn't land on a particular face.} and, consequently, the power set is the $\sigma$-algebra.
@@
 
-\textbf{Absolute continuity of probability measures.} As we saw before, probability measures over a single point over the continuum cannot be understood as valid ensembles.
+\textbf{Absolute continuity of probability measures.} As we saw before, probability measures over a single point on the continuum cannot be understood as valid ensembles.
 To address this, we add a requirement that is both physically meaningful and mathematically useful.
@@
 This measure is additive over disjoint regions as they represent mutually exclusive events.
-The requirement is that we can assign non-zero probability only to events that have a non-zero count of states.
+The requirement is that we can assign a non-zero probability only to events that have a non-zero count of states.
 That is, if for a particular event $A$ we have $p(A) \neq 0$, then we must have that $\mu(A) \neq 0$ as well.
@@
 The core insight from Chouqet theory, then, is that both in the classical and quantum case we can represent ensembles as a probability distribution over a set of states.
-However, only in the classical case this representation is unique.
+However, only in the classical case is this representation unique.
 In the quantum case, mixed states can be decomposed into pure states in multiple ways, and it is this multiple decomposition that makes the ensemble space not a simplex, and therefore the space not classical.
@@
 
-\textbf{Probability of outcomes vs fraction of components.} We saw that classical probability fails in quantum mechanics because pure states are not always mutually exclusive, are not orthogonal.
+\textbf{Probability of outcomes vs fraction of components.} We saw that classical probability fails in quantum mechanics because pure states are not always mutually exclusive, i.e.
+are not orthogonal.
 That is why, in quantum mechanics, probability can be defined only once a context, an orthogonal basis, is chosen.
@@
 
-Note that the linearity in our axioms is in terms of mixing coefficients, not probability of outcomes.
+Note that the linearity in our axioms is in terms of mixing coefficients, not the probability of outcomes.
 That is, regardless of whether we can characterize experiment outcomes with probability distributions, preparation procedures will always allow us to create a statistical mixture.
@@
 If we take a set $A$, we can ask: what is the highest entropy reachable by a mixture of elements of $A$?
-If $A$ happens to be the set of ensembles with support $U$, then the supremum of the entropy is $\log \mu(U)$, which works also in the unbounded case.
+If $A$ happens to be the set of ensembles with support $U$, then the supremum of the entropy is $\log \mu(U)$, which also works in the unbounded case.
 
@@
 In the classical case, disjoint sets of extreme points will be orthogonal, meaning that the state capacity there restricted is an additive measure.
-In the discrete case, every point is associated to zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
+In the discrete case, every point is associated with zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
 In the continuous case, gives a subset of phase space $U$ and $A_U$ the set of distributions with support within $U$, the maximum entropy is reached by the uniform distribution $\rho_U \in A_U$, which means we recover $\log_2 \capacity(A_U) = S(\rho_U)$, the fundamental postulate of statistical mechanics.
@@
 The state capacity, then, recovers the number of distinguishable states.
-In classical mechanics, all states are perfectly distinguishable as they represent mutually exclusive cases, the measure is additive.
+In classical mechanics, all states are perfectly distinguishable as they represent mutually exclusive cases, and the measure is additive.
 In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states.
@@
 \textbf{Requirements for non-additive probability theory.} In a previous section we reviewed what types of objects we have in probability theory for classical ensembles: a sample space $X$, a probability measure $p$, random variables $F : X \to \mathbb{R}$, expectations as integrals of the probability measure $E[F] = \int_X F(x) dp$, a state counting measure $\mu$, the probability density as the Radon-Nikodym derivative $\frac{dp}{d\mu}$ and the entropy as $\int_X \rho \log \rho d\mu$.
-Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the states counting measure $\mu$, is there a generalization of all other elements?
+Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of the probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the states counting measure $\mu$, is there a generalization of all other elements?
 That is, can we find an integral and derivative operation over non-additive measures that generalizes to all physical theories?
@@
 \end{equation}
-This last requirement may not be possible, or it may be used as a guide in case that the previous derivative is not uniquely defined.
+This last requirement may not be possible, or it may be used as a guide in case the previous derivative is not uniquely defined.
 

### 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: grammar
Three more small grammatical fixes.

@@
 As part of our larger project, we are developing a general theory of ensemble spaces, the full treatment of which is beyond the scope of this work.
-We limit ourselves to give the initial premises and the final mathematical characterization, without justifications and proofs.\footnote{More details are available in our ongoing work (cite)}
+We limit ourselves to giving the initial premises and the final mathematical characterization, without justifications and proofs.\footnote{More details are available in our ongoing work (cite)}
 
@@
 
-The above theorem tells us that the map between probability measures and ensembles is surjective. The second question, then, when is it bijective? Again, Choquet theory has a clear answer.
+The above theorem tells us that the map between probability measures and ensembles is surjective. The second question, then, is when is it bijective? Again, Choquet theory has a clear answer.
 \begin{thrm}[Choquet 2]
@@
 We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutual exclusiveness, is responsible for the failure of the additivity of the measure.
-Note that, since the fraction capacity depends only on the ensemble space, and not also on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
+Note that, since the fraction capacity depends only on the ensemble space, and not on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
 

### 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Second Choquet integral figure
This TikZ code reproduces the second sketch for the Choquet integral calculation.

--- 2025-07-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 Therefore, at a conceptual level, probabilities can only give us a patchwork representation of ensembles.
-Some authors embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.
+Some author embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.
 [CITATION NEEDED] However, there is a different route.
@@
 In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states.
-We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutual exclusiveness, is responsible for the failure of the additivity of the measure.
+We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutually exclusiveness, is responsible for the failure of the additivity of the measure.
 Note that, since the fraction capacity depends only on the ensemble space, and not also on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{ChoquetCalculation2.jpg}
+
+	% perspective
+	\newcommand{\yaw}{11}
+	\newcommand{\pitch}{28}
+	\newcommand{\roll}{23}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	% dimensions
+	\newcommand{\s}{0.5}
+	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
+	\pgfmathsetmacro{\stheta}{asin(\s)}
+	\pgfmathsetmacro{\minstheta}{asin(-\s)}
+
+	\newcommand{\radius}{1}
+	\newcommand{\margin}{1.3}
+
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
+		% Bloch sphere (outline drawn in canvas plane)
+		\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, radius=1 cm] coordinate (A);
+%		\draw[dash dot, red] (A) arc[start angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, end angle=-90, radius=1 cm];
+		\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, radius=1 cm] coordinate (B);
+%		\draw[dash dot, green] (B) arc[start angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, end angle=270, radius=1 cm];
+		\draw[dashed] (A) arc[start angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)-360}, radius=1 cm];
+
+		% special points
+		\shadedraw (0, 0, 0) circle (0.03 cm) node[below right] {$\mathsf{o}$};
+		\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
+		\shadedraw (0, -\radius, 0) circle (0.03cm) node[below] {$\mathsf{a}$};
+		\shadedraw (0, \s, 0) circle (0.03cm) node[left] {$\mathsf{e}$};
+
+		% convex combination
+		\draw (0, -\radius, 0) -- (0, \s, 0);
+		\draw (2.25*\xx, 2.25*\yx, 2.25*\zx) node {$\ens[o] = \frac{\overline{\ens[o]\ens[a]}}{\overline{\ens[e]\ens[a]}} \ens[e] + \frac{\overline{\ens[o]\ens[e]}}{\overline{\ens[e]\ens[a]}} \ens[a]$};
+
+		% s axis
+		\begin{scope}[shift={(-1.5*\xx, -1.5*\yx, -1.5*\zx)}]
+			\draw (0, -1, 0) -- (0, 0, 0);
+			\draw[->] (0, 0, 0) -- (0, 1, 0) node[below left] {$s$};
+		\end{scope}
+
+		% cut
+		\begin{scope}[canvas is zx plane at y=\s]
+%			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+		\end{scope}
+
+		% meridians and equator
+		\begin{scope}[canvas is yz plane at x=0]
+			% foreground above s
+			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=90-\stheta, radius=1 cm] coordinate (C);
+			% foreground below s
+			\draw[dashed] (C) arc[start angle=90-\stheta, end angle={atan2(-\ixz, -\ixy)}, radius=1 cm] coordinate (D);
+			% background below s
+			\draw[loosely dotted] (D) arc[start angle={atan2(-\ixz, -\ixy)}, end angle=\stheta+270, radius=1 cm];
+			% background above s
+			\draw[densely dotted] (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=\stheta-90, radius=1 cm];
+		\end{scope}
+%		\begin{scope}[canvas is xy plane at z=0]
+%			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+%			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+%		\end{scope}
+		\begin{scope}[canvas is xy plane at z=0]
+			% foreground above s
+			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=180-\stheta, radius=1 cm] coordinate (E);
+			% foreground below s
+			\draw[dashed] (E) arc[start angle=180-\stheta, end angle={atan2(\izy, \izx)+360}, radius=1 cm] coordinate (F);
+			% background below s
+			\draw[loosely dotted] (F) arc[start angle={atan2(\izy, \izx)}, end angle=\stheta, radius=1 cm];
+			% background above s
+			\draw[densely dotted] (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=\stheta, radius=1 cm];
+		\end{scope}
+	\end{tikzpicture}
 	\caption{...}

### 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: caption for second Choquet integral figure
NonAdditiveProbability: Choquet integral plot
I added a python script that draws a plot of the integrand fcap_o(A(s)) as a function of s and included the result in a figure.

@@
 	\centering
-
-	% perspective
-	\newcommand{\yaw}{11}
-	\newcommand{\pitch}{28}
-	\newcommand{\roll}{23}
-
-	% transformation matrix = components of data axes in canvas coordinates
-	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
-	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
-	\pgfmathsetmacro{\xz}{-sin(\pitch)}
-
-	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
-	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
-	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
-
-	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
-	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
-	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
-
-	% intersection of yz grand circle with horizon
-	\pgfmathsetmacro{\ixx}{0}
-	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
-	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
-	% intersection of zx grand circle with horizon
-	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
-	\pgfmathsetmacro{\iyy}{0}
-	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
-	% intersection of xy grand circle with horizon
-	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
-	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
-	\pgfmathsetmacro{\izz}{0}
-
-	% dimensions
-	\newcommand{\s}{0.5}
-	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
-	\pgfmathsetmacro{\stheta}{asin(\s)}
-	\pgfmathsetmacro{\minstheta}{asin(-\s)}
-
-	\newcommand{\radius}{1}
-	\newcommand{\margin}{1.3}
-
-	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
-		% Bloch sphere (outline drawn in canvas plane)
-		\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, radius=1 cm] coordinate (A);
-%		\draw[dash dot, red] (A) arc[start angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, end angle=-90, radius=1 cm];
-		\draw (\xy, \yy, \zy) arc[start angle=90, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, radius=1 cm] coordinate (B);
-%		\draw[dash dot, green] (B) arc[start angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)}, end angle=270, radius=1 cm];
-		\draw[dashed] (A) arc[start angle={acos(\radiusats*\iyx*\xx - \s*\yx + \radiusats*\iyz*\zx)}, end angle={acos(-\radiusats*\iyx*\xx - \s*\yx - \radiusats*\iyz*\zx)-360}, radius=1 cm];
-
-		% special points
-		\shadedraw (0, 0, 0) circle (0.03 cm) node[below right] {$\mathsf{o}$};
-		\shadedraw (0, \radius, 0) circle (0.03 cm) node[above] {$\mathsf{b}$};
-		\shadedraw (0, -\radius, 0) circle (0.03cm) node[below] {$\mathsf{a}$};
-		\shadedraw (0, \s, 0) circle (0.03cm) node[left] {$\mathsf{e}$};
-
-		% convex combination
-		\draw (0, -\radius, 0) -- (0, \s, 0);
-		\draw (2.25*\xx, 2.25*\yx, 2.25*\zx) node {$\ens[o] = \frac{\overline{\ens[o]\ens[a]}}{\overline{\ens[e]\ens[a]}} \ens[e] + \frac{\overline{\ens[o]\ens[e]}}{\overline{\ens[e]\ens[a]}} \ens[a]$};
-
-		% s axis
-		\begin{scope}[shift={(-1.5*\xx, -1.5*\yx, -1.5*\zx)}]
-			\draw (0, -1, 0) -- (0, 0, 0);
-			\draw[->] (0, 0, 0) -- (0, 1, 0) node[below left] {$s$};
-		\end{scope}
-
-		% cut
-		\begin{scope}[canvas is zx plane at y=\s]
-%			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
-			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
-		\end{scope}
-
-		% meridians and equator
-		\begin{scope}[canvas is yz plane at x=0]
-			% foreground above s
-			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=90-\stheta, radius=1 cm] coordinate (C);
-			% foreground below s
-			\draw[dashed] (C) arc[start angle=90-\stheta, end angle={atan2(-\ixz, -\ixy)}, radius=1 cm] coordinate (D);
-			% background below s
-			\draw[loosely dotted] (D) arc[start angle={atan2(-\ixz, -\ixy)}, end angle=\stheta+270, radius=1 cm];
-			% background above s
-			\draw[densely dotted] (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, end angle=\stheta-90, radius=1 cm];
-		\end{scope}
-%		\begin{scope}[canvas is xy plane at z=0]
-%			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
-%			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
-%		\end{scope}
-		\begin{scope}[canvas is xy plane at z=0]
-			% foreground above s
-			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=180-\stheta, radius=1 cm] coordinate (E);
-			% foreground below s
-			\draw[dashed] (E) arc[start angle=180-\stheta, end angle={atan2(\izy, \izx)+360}, radius=1 cm] coordinate (F);
-			% background below s
-			\draw[loosely dotted] (F) arc[start angle={atan2(\izy, \izx)}, end angle=\stheta, radius=1 cm];
-			% background above s
-			\draw[densely dotted] (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, end angle=\stheta, radius=1 cm];
-		\end{scope}
-	\end{tikzpicture}
+	\includegraphics[width=0.8\columnwidth]{ChoquetCalculation2.jpg}
 	\caption{...}
@@
 \end{equation}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.4\textwidth]{ChoquetPlot.png}
+	\caption{The integrand $\frcap_{\ens[o]}(A(s))$ as a function of $s$.}
+\end{figure}
 We can now calculate the integral

### 2025-07-30 Articles/2024-NonAdditiveProbability/ChoquetPlot.py

@@ -1,0 +1,52 @@
+#!/usr/bin/env python3
+""" Choquet integral example in NonAdditiveProbability
+
+Created: Wed Jul 30, 2025, 12:16 PM EST
+Author: Tobias Thrien
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+from matplotlib.ticker import NullFormatter, FixedLocator, FixedFormatter
+
+a = 0
+b = 1
+margin = 0.05 * (b - a)
+s_min, s_max = a - margin, b + margin
+
+
+def fcap(s):
+    s = np.array(s)
+    return np.where(s > b, 0,
+            np.where(s <= 1/2*(a+b), 1,
+                1/2*(b-a)/(s-a)))
+
+S = np.linspace(s_min, s_max, endpoint=True, num=300)
+F = fcap(S)
+
+fig, ax = plt.subplots()
+
+# TODO mark discontinuity by splitting array above b
+ax.plot(S, F, linewidth=3)
+
+ax.set_xlabel("$s$")
+ax.xaxis.set_minor_formatter(NullFormatter())
+ax.xaxis.set_major_locator(FixedLocator([a, 1/2*(a+b), b]))
+ax.xaxis.set_major_formatter(FixedFormatter(
+    ["$a$", r"${}^1/_2(a+b)$", "$b$"]))
+ax.set_xlim(s_min, s_max)
+
+ax.set_ylabel("$\mathrm{fcap}_{\mathsf{o}}(A(s))$")
+ax.set_ylim(bottom=0)
+y_min, y_max = ax.get_ylim()
+
+ax.spines[["top", "right"]].set_visible(False)
+ax.plot(s_max, y_min, ">k", clip_on=False)
+ax.plot(s_min, y_max, "^k", clip_on=False)
+# This doesn't work for some reason
+#ax.plot(1, 0, ">k", transform=ax.get_yaxis_transform(), clip_on=False)
+# The pointer on the y axis appears above a instead of s_min
+#ax.plot(0, 1, "^k", transform=ax.get_xaxis_transform(), clip_on=False)
+
+plt.tight_layout()
+fig.savefig("ChoquetPlot.png")
+plt.close(fig)

### 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Figure with ensemble spaces
Added a figure with examples of ensemble spaces:
 - classical system with 3 points (triangle)
 - 2D quantum system (Bloch ball)

--- 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-TODO: figure with classical triangle and quantum Bloch Ball
+\begin{figure}[H]
+	\centering
+
+	% perspective
+	\newcommand{\yaw}{11}
+	\newcommand{\pitch}{28}
+	\newcommand{\roll}{23}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	% dimensions
+	\newcommand{\radius}{1}
+
+	\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
+		% Triangle
+		\draw (canvas cs: x=-3cm, y=-0.25cm) +(canvas polar cs: angle=90, radius=1cm) -- +(canvas polar cs: angle=210, radius=1cm) -- +(canvas polar cs: angle=330, radius=1cm) -- cycle;
+
+		% Bloch sphere (outline drawn in canvas plane)
+		\draw (0, 0, 0) circle (\radius cm);
+
+		% meridians and equator
+		\begin{scope}[canvas is yz plane at x=0]
+			\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+			\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is zx plane at y=0]
+			\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+			\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+		\end{scope}
+		\begin{scope}[canvas is xy plane at z=0]
+			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+		\end{scope}
+	\end{tikzpicture}
+	\caption{Examples of a classical (triangle) and quantum (ball) ensemble space in three and two dimensions, respectively.}
+\end{figure}
 
@@
 	\end{tikzpicture}
-	\caption{...}
+	\caption{Geometrical representation of the element $\ens[e]$ in $A(s)$ with the largest fraction of $\ens[o]$.}
 \end{figure}

### 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Figure with ensemble spaces
Added a figure with examples of ensemble spaces:
 - classical system with 3 points (triangle)
 - 2D quantum system (Bloch ball)

--- 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 \end{equation}
-\begin{figure}[H]
-	\centering
-	\includegraphics[width=0.4\textwidth]{ChoquetPlot.png}
-	\caption{The integrand $\frcap_{\ens[o]}(A(s))$ as a function of $s$.}
-\end{figure}
 We can now calculate the integral

### 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Plot Choquet integrand in TikZ
Replace the matplotlib code with TikZ code.

--- 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{ChoquetPlot.png}
+	\newcommand{\ticklength}{0.03}
+	\begin{tikzpicture}[scale=3]
+		% axes
+		\draw[->] (0, 0) -- (1.3, 0);
+		\draw (0, 0) node[left] {$0$} -- +(-\ticklength, 0) (0, 0.5) node[left] {$1/2$} -- +(-\ticklength, 0) (0, 1) node[left] {$1$} -- +(-\ticklength, 0);
+		\draw[->] (0, 0) -- (0, 1.1);
+		\draw (0, 0) node[below] {$a$} -- +(0, -\ticklength) (0.5, 0) node[below] {$1/2(a+b)$} -- +(0, -\ticklength) (1, 0) node[below] {$b$} -- +(0, -\ticklength);
+		% cases s <= 1/2(a+b) and 1/2(a+b) < s <= b
+		\draw[thick, blue, domain=0.5:1, samples=200] (0, 1) --plot (\x, {0.5/\x});
+		% mark discontinuity
+		\filldraw (1, 0.5) circle (0.02cm);
+		\draw (1, 0) circle (0.02cm);
+		% case s > b
+		\draw[thick, blue] (1, 0) -- (1.2, 0);
+	\end{tikzpicture}
 	\caption{The integrand $\frcap_{\ens[o]}(A(s))$ as a function of $s$.}

### 2025-07-30 Articles/2024-NonAdditiveProbability/ChoquetPlot.py

@@ -1,52 +1,0 @@
-#!/usr/bin/env python3
-""" Choquet integral example in NonAdditiveProbability
-
-Created: Wed Jul 30, 2025, 12:16 PM EST
-Author: Tobias Thrien
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-from matplotlib.ticker import NullFormatter, FixedLocator, FixedFormatter
-
-a = 0
-b = 1
-margin = 0.05 * (b - a)
-s_min, s_max = a - margin, b + margin
-
-
-def fcap(s):
-    s = np.array(s)
-    return np.where(s > b, 0,
-            np.where(s <= 1/2*(a+b), 1,
-                1/2*(b-a)/(s-a)))
-
-S = np.linspace(s_min, s_max, endpoint=True, num=300)
-F = fcap(S)
-
-fig, ax = plt.subplots()
-
-# TODO mark discontinuity by splitting array above b
-ax.plot(S, F, linewidth=3)
-
-ax.set_xlabel("$s$")
-ax.xaxis.set_minor_formatter(NullFormatter())
-ax.xaxis.set_major_locator(FixedLocator([a, 1/2*(a+b), b]))
-ax.xaxis.set_major_formatter(FixedFormatter(
-    ["$a$", r"${}^1/_2(a+b)$", "$b$"]))
-ax.set_xlim(s_min, s_max)
-
-ax.set_ylabel("$\mathrm{fcap}_{\mathsf{o}}(A(s))$")
-ax.set_ylim(bottom=0)
-y_min, y_max = ax.get_ylim()
-
-ax.spines[["top", "right"]].set_visible(False)
-ax.plot(s_max, y_min, ">k", clip_on=False)
-ax.plot(s_min, y_max, "^k", clip_on=False)
-# This doesn't work for some reason
-#ax.plot(1, 0, ">k", transform=ax.get_yaxis_transform(), clip_on=False)
-# The pointer on the y axis appears above a instead of s_min
-#ax.plot(0, 1, "^k", transform=ax.get_xaxis_transform(), clip_on=False)
-
-plt.tight_layout()
-fig.savefig("ChoquetPlot.png")
-plt.close(fig)

### 2025-07-31 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: minor changes to captions

--- 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 	\end{tikzpicture}
-	\caption{Examples of a classical (triangle) and quantum (ball) ensemble space in three and two dimensions, respectively.}
+	\caption{Examples of a classical discrete ensemble space with three pure states (triangle) and of a two dimensional quantum ensemble space (ball).}
 \end{figure}
@@
 	\end{tikzpicture}
-	\caption{Geometrical representation of the element $\ens[e]$ in $A(s)$ with the largest fraction of $\ens[o]$.}
+	\caption{Geometric representation of the element $\ens[e]$ in $A(s)$ with the largest fraction of $\ens[o]$.}
 \end{figure}

### 2025-07-31 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Include 0 in Choquet plot
I added a zero to indicate that a and b are both positive in our example.
I also adjusted the positioning of the tick labels.

--- 2025-07-30 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 	\newcommand{\ticklength}{0.03}
-	\begin{tikzpicture}[scale=3]
+	\begin{tikzpicture}[scale=3, xticklabel/.style={anchor=base, inner sep=1pt, yshift=-2.5ex}]
 		% axes
-		\draw[->] (0, 0) -- (1.3, 0);
+		\draw[->] (0, 0) -- (1.7, 0);
 		\draw (0, 0) node[left] {$0$} -- +(-\ticklength, 0) (0, 0.5) node[left] {$1/2$} -- +(-\ticklength, 0) (0, 1) node[left] {$1$} -- +(-\ticklength, 0);
-		\draw[->] (0, 0) -- (0, 1.1);
-		\draw (0, 0) node[below] {$a$} -- +(0, -\ticklength) (0.5, 0) node[below] {$1/2(a+b)$} -- +(0, -\ticklength) (1, 0) node[below] {$b$} -- +(0, -\ticklength);
+		\draw[->] (0, 0) -- (0, 1.2);
+		\draw (0, 0) node[xticklabel] {$0$} -- +(0, -\ticklength) (0.5, 0) node[xticklabel] {$a$} -- +(0, -\ticklength) (1, 0) node[xticklabel] {$1/2(a+b)$} -- +(0, -\ticklength) (1.5, 0) node[xticklabel] {$b$} -- +(0, -\ticklength);
 		% cases s <= 1/2(a+b) and 1/2(a+b) < s <= b
-		\draw[thick, blue, domain=0.5:1, samples=200] (0, 1) --plot (\x, {0.5/\x});
+		\draw[thick, blue, domain=1:1.5, samples=200] (0, 1) --plot (\x, {0.5/(\x-0.5)});
 		% mark discontinuity
-		\filldraw (1, 0.5) circle (0.02cm);
-		\draw (1, 0) circle (0.02cm);
+		\filldraw (1.5, 0.5) circle (0.02cm);
+		\draw (1.5, 0) circle (0.02cm);
 		% case s > b
-		\draw[thick, blue] (1, 0) -- (1.2, 0);
+		\draw[thick, blue] (1.5, 0) -- (1.7, 0);
 	\end{tikzpicture}

### 2025-07-31 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Squashed commit of the following:
commit 3af5bed94c6d67ad422763a02400c9575b2c1ba7
Author: thrien <39382079+thrien@users.noreply.github.com>
Date:   Thu Jul 31 15:38:59 2025 -0400

    NonAdditiveProbability: Include 0 in Choquet plot

    I added a zero to indicate that a and b are both positive in our example.
    I also adjusted the positioning of the tick labels.

commit 7cebf7194eb2c404d12e1090c3905d5c00d1df9b
Author: thrien <39382079+thrien@users.noreply.github.com>
Date:   Wed Jul 30 17:02:51 2025 -0400

    NonAdditiveProbability: Plot Choquet integrand in TikZ

    Replace the matplotl

--- 2025-07-31 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 \end{equation}
+\begin{figure}[H]
+	\centering
+	\newcommand{\ticklength}{0.03}
+	\begin{tikzpicture}[scale=3, xticklabel/.style={anchor=base, inner sep=1pt, yshift=-2.5ex}]
+		% axes
+		\draw[->] (0, 0) -- (1.7, 0);
+		\draw (0, 0) node[left] {$0$} -- +(-\ticklength, 0) (0, 0.5) node[left] {$1/2$} -- +(-\ticklength, 0) (0, 1) node[left] {$1$} -- +(-\ticklength, 0);
+		\draw[->] (0, 0) -- (0, 1.2);
+		\draw (0, 0) node[xticklabel] {$0$} -- +(0, -\ticklength) (0.5, 0) node[xticklabel] {$a$} -- +(0, -\ticklength) (1, 0) node[xticklabel] {$1/2(a+b)$} -- +(0, -\ticklength) (1.5, 0) node[xticklabel] {$b$} -- +(0, -\ticklength);
+		% cases s <= 1/2(a+b) and 1/2(a+b) < s <= b
+		\draw[thick, blue, domain=1:1.5, samples=200] (0, 1) --plot (\x, {0.5/(\x-0.5)});
+		% mark discontinuity
+		\filldraw (1.5, 0.5) circle (0.02cm);
+		\draw (1.5, 0) circle (0.02cm);
+		% case s > b
+		\draw[thick, blue] (1.5, 0) -- (1.7, 0);
+	\end{tikzpicture}
+	\caption{The integrand $\frcap_{\ens[o]}(A(s))$ as a function of $s$.}
+\end{figure}
 We can now calculate the integral

### 2025-07-31 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: halved height of the graph
NonAdditiveProbability: grammar (modified from Tobias)

@@
 		\draw[->] (0, 0) -- (1.7, 0);
-		\draw (0, 0) node[left] {$0$} -- +(-\ticklength, 0) (0, 0.5) node[left] {$1/2$} -- +(-\ticklength, 0) (0, 1) node[left] {$1$} -- +(-\ticklength, 0);
-		\draw[->] (0, 0) -- (0, 1.2);
+		\draw (0, 0) node[left] {$0$} -- +(-\ticklength, 0) (0, 0.25) node[left] {$1/2$} -- +(-\ticklength, 0) (0, 0.5) node[left] {$1$} -- +(-\ticklength, 0);
+		\draw[->] (0, 0) -- (0, 0.6);
 		\draw (0, 0) node[xticklabel] {$0$} -- +(0, -\ticklength) (0.5, 0) node[xticklabel] {$a$} -- +(0, -\ticklength) (1, 0) node[xticklabel] {$1/2(a+b)$} -- +(0, -\ticklength) (1.5, 0) node[xticklabel] {$b$} -- +(0, -\ticklength);
 		% cases s <= 1/2(a+b) and 1/2(a+b) < s <= b
-		\draw[thick, blue, domain=1:1.5, samples=200] (0, 1) --plot (\x, {0.5/(\x-0.5)});
+		\draw[thick, blue, domain=1:1.5, samples=200] (0, 0.5) --plot (\x, {0.25/(\x-0.5)});
 		% mark discontinuity
-		\filldraw (1.5, 0.5) circle (0.02cm);
+		\filldraw (1.5, 0.25) circle (0.02cm);
 		\draw (1.5, 0) circle (0.02cm);

### 2025-08-01 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: minor cleanup
Deleted and modified unnecessary comments from my TikZ code.
NonAdditiveProbability: fixed jumbled introduction

@@
 
-		% meridians and equator
+		% meridians
 		\begin{scope}[canvas is yz plane at x=0]
@@
 		\end{scope}
-%		\begin{scope}[canvas is xy plane at z=0]
-%			\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
-%			\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
-%		\end{scope}
 		\begin{scope}[canvas is xy plane at z=0]

### 2025-08-01 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: SIC-POVM figure
Added a figure to illustrate a SIC-POVM on the Bloch ball, which is a tetrahedron inside the ball.
Alternatively, one could define a tetrahedron that fits the ball inside, which is illustrated in a separate subfigure.
There is not enough space on the figure to label all the extreme points as was done here:
https://commons.wikimedia.org/wiki/File:Regular_tetrahedron_inscribed_in_a_sphere.svg
The text would be too small.
I left the labels as comments anyway.

@@
 Since the map is linear, and expectation values are linear, one can express expectation values as integrals or sums over the quasi-probability $w(z)$.
+
+\begin{figure}[H]
+	\centering
+
+	% perspective
+	% from scipy.spatial.transform import Rotation as R
+	% R.from_euler("yx", [50, 10], degrees=True).as_euler("xyz", degrees=True).round(0)[::-1]
+	\newcommand{\yaw}{12}
+	\newcommand{\pitch}{49}
+	\newcommand{\roll}{15}
+
+	% transformation matrix = components of data axes in canvas coordinates
+	\pgfmathsetmacro{\xx}{cos(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xy}{sin(\yaw)*cos(\pitch)}
+	\pgfmathsetmacro{\xz}{-sin(\pitch)}
+
+	\pgfmathsetmacro{\yx}{cos(\yaw)*sin(\pitch)*sin(\roll) - sin(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yy}{sin(\yaw)*sin(\pitch)*sin(\roll) + cos(\yaw)*cos(\roll)}
+	\pgfmathsetmacro{\yz}{cos(\pitch)*sin(\roll)}
+
+	\pgfmathsetmacro{\zx}{cos(\yaw)*sin(\pitch)*cos(\roll) + sin(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zy}{sin(\yaw)*sin(\pitch)*cos(\roll) - cos(\yaw)*sin(\roll)}
+	\pgfmathsetmacro{\zz}{cos(\pitch)*cos(\roll)}
+
+	% intersection of yz grand circle with horizon
+	\pgfmathsetmacro{\ixx}{0}
+	\pgfmathsetmacro{\ixy}{\zz / sqrt(\zz*\zz + \yz*\yz)}
+	\pgfmathsetmacro{\ixz}{-\yz / sqrt(\zz*\zz + \yz*\yz)}
+	% intersection of zx grand circle with horizon
+	\pgfmathsetmacro{\iyx}{\zz / sqrt(\zz*\zz + \xz*\xz)}
+	\pgfmathsetmacro{\iyy}{0}
+	\pgfmathsetmacro{\iyz}{-\xz / sqrt(\zz*\zz + \xz*\xz)}
+	% intersection of xy grand circle with horizon
+	\pgfmathsetmacro{\izx}{-\yz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izy}{\xz / sqrt(\xz*\xz + \yz*\yz)}
+	\pgfmathsetmacro{\izz}{0}
+
+	\newcommand{\radius}{1}
+
+	\begin{subfigure}[c]{0.23\textwidth}
+		\centering
+		\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=1.2]
+			% Bloch sphere (outline drawn in canvas plane)
+			\draw (0, 0, 0) circle (\radius cm);
+
+			% meridians and equator
+			\begin{scope}[canvas is yz plane at x=0]
+				\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+				\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is zx plane at y=0]
+				\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+				\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is xy plane at z=0]
+				\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+				\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+			\end{scope}
+
+			% tetrahedron
+			\coordinate (A) at (0, 1, 0);
+			\coordinate (B) at (0, -1/3, {sqrt(8/9)});
+			\coordinate (C) at ({sqrt(2/3)}, -1/3, {-sqrt(2/9)});
+			\coordinate (D) at ({-sqrt(2/3)}, -1/3, {-sqrt(2/9)});
+			\begin{scope}[every node/.style={font=\tiny}]
+				\draw (A) -- (B) -- (C) -- cycle;
+				\draw (A) -- (B) -- (D) -- cycle;
+				\draw (A) -- (C) -- (D) -- cycle;
+				\draw (B) -- (C) -- (D) -- cycle;
+%				\shadedraw (A) circle (0.03 cm) node[above] {$| 0 \rangle$};
+%				\shadedraw (B) circle (0.03 cm) node[below] {$\frac{1}{\sqrt{3}} | 0 \rangle + \sqrt{\frac{2}{3}} e^{i\frac{2\pi}{3}} | 1 \rangle$};
+%				\shadedraw (C) circle (0.03 cm) node[above] {$\frac{1}{\sqrt{3}} | 0 \rangle + \sqrt{\frac{2}{3}} e^{i\frac{4\pi}{3}} | 1 \rangle$};
+%				\shadedraw (D) circle (0.03 cm) node[below] {$\frac{1}{\sqrt{3}} | 0 \rangle + \sqrt{\frac{2}{3}} | 1 \rangle$};
+			\end{scope}
+		\end{tikzpicture}
+		\caption{}
+	\end{subfigure}
+	\begin{subfigure}[c]{0.23\textwidth}
+		\centering
+		\begin{tikzpicture}[x={(\xx cm, \xy cm)}, y={(\yx cm, \yy cm)}, z={(\zx cm, \zy cm)}, scale=0.65]
+			% Bloch sphere (outline drawn in canvas plane)
+			\draw (0, 0, 0) circle (\radius cm);
+
+			% meridians and equator
+			\begin{scope}[canvas is yz plane at x=0]
+				\draw (\ixy, \ixz) arc[start angle={atan2(\ixz, \ixy)}, delta angle=180, radius=1 cm];
+				\draw[dashed] (-\ixy, -\ixz) arc[start angle={atan2(-\ixz, -\ixy)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is zx plane at y=0]
+				\draw[dashed] (\iyz, \iyx) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=1 cm];
+				\draw (-\iyz, -\iyx) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=1 cm];
+			\end{scope}
+			\begin{scope}[canvas is xy plane at z=0]
+				\draw[dashed] (\izx, \izy) arc[start angle={atan2(\izy, \izx)}, delta angle=180, radius=1 cm];
+				\draw (-\izx, -\izy) arc[start angle={atan2(-\izy, -\izx)}, delta angle=180, radius=1 cm];
+			\end{scope}
+
+			% inscribed tetrahedron
+			\begin{scope}[scale=3, every node/.style={font=\small}]
+				\coordinate (A) at (0, 1, 0);
+				\coordinate (B) at (0, -1/3, {sqrt(8/9)});
+				\coordinate (C) at ({sqrt(2/3)}, -1/3, {-sqrt(2/9)});
+				\coordinate (D) at ({-sqrt(2/3)}, -1/3, {-sqrt(2/9)});
+				\draw (A) -- (B) -- (C) -- cycle;
+				\draw (A) -- (B) -- (D) -- cycle;
+				\draw (A) -- (C) -- (D) -- cycle;
+				\draw (B) -- (C) -- (D) -- cycle;
+			\end{scope}
+		\end{tikzpicture}
+		\caption{}
+	\end{subfigure}
+	\caption{(a) Example of a SIC-POVM on a Bloch ball.
+The four extreme points of the tetrahedron represent the states $P_i$.
+States with positive $\lambda_i$ lie within the tetrahedron, while all other states require negative coefficients.
+(b) The Bloch ball can alternatively be circumscribed by a larger tetrahedron to guarantee positive coefficients, but the extreme points are no longer physical states.}
+\end{figure}
 

### 2025-08-05 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: fixed caption

@@
 States with positive $\lambda_i$ lie within the tetrahedron, while all other states require negative coefficients.
-(b) The Bloch ball can alternatively be circumscribed by a larger tetrahedron to guarantee positive coefficients, but the extreme points are no longer physical states.}
+(b) The Bloch ball can alternatively be inscribed within a larger tetrahedron to guarantee positive coefficients, but the extreme points are no longer physical states.}
 \end{figure}

### 2025-08-07 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Illustrated dice distributions
I illustrated the distributions on a dice as it was done in the book.

This first draft uses tikz pictures as symbols in an equation.

@@
 Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
-Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$, where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/6$ of $\ens$ can be constructed from $A$, but no more.
+Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$ (see \eqref{eq:dice} for an illustration), where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/5$ of $\ens$ can be constructed from $A$, but no more.
 Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
-Note that, given the uniform distribution, $1/6$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
+Note that, given the uniform distribution, $1/5$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
 This gives the basic insight for our definitions.
+
+\begin{equation}\label{eq:dice}
+	\def\xstep{0.75em}
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
+	\end{tikzpicture}
+	= \frac{1}{6}
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1)};
+	\end{tikzpicture}
+	+ \frac{5}{6}
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
+	\end{tikzpicture}
+\end{equation}
 

### 2025-08-08 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed introduction

--- 2025-08-05 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 The starting point is that any physical law deals with experimentally reproducible relationships.
-This implies a physical theory must gives us the space of statistical ensembles (i.e.
+This implies a physical theory must give us the space of statistical ensembles (i.e.
 the collection of all possible preparation procedures) that the theory allows.
-As a basic requirements, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from GPTs is the entropy requirement.
-We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
-Mathematically, under these premises, an ensembles space is a convex subset of a locally convex topological vector space.
+As a basic requirement, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from GPTs is the entropy requirement.
+We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough real valued quantities to identify each ensemble (i.e.
+there is a family of affine functions that identify each ensemble through their values).
+Mathematically, these premises lead us to assume that an ensemble space is a convex subset of a metrizable locally convex topological vector space.
 The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the dimension of orthogonal subspaces of an operator)?
@@
 
-We will first review how measure theory is used in classical probability together with some results from Choquet theory.
-This will tell us that elements of a compact convex set can always be represented by a probability measure over its extreme points, the pure states.
+We will first review how measure theory is used in classical probability and classical statistical mechanics.
+We will connect this to Choquet theory, which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.
+each ensemble can be represented by a probability measure over pure states).
 While such representation is always possible, at least in the compact case, it is not unique.
-Uniqueness requires the space to be a Choquet simplex, which is a generalization of the standard finite-dimensional simplex and represents the classical case.
-The key insight, then, an ensemble space is classical if and only if each ensembles admits a unique decomposition in terms of pure states.
+Uniqueness requires the space to be a Choquet simplex, the generalization of the finite-dimensional simplex, and represents the classical case.
+The key insight, then, is that an ensemble space is classical if and only if each ensembles admits a unique decomposition in terms of pure states.
 
 We will then review some results from the use of quasi-probability measures.
+The most used is the Wigner function, that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through normal integral.
 While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.
-The key insight is that it is not negative probability that characterizes quasi-probability, but the lack of mutual exclusivity of the points of the space.
+The key insight is that it is not negative probability that characterizes quasi-probability,\footnote{Some representation guarantee the quasi-probability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
 The quasi-probability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
@@
 The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
-Given an ensemble $\ens$, we can ask what fraction can be produces by a mixture of a given ensembles.
+Given an ensemble $\ens$, we can ask what fraction can be produced by a mixture of a given ensembles.
 This leads to the fraction capacity: a sub-additive set function that, like the probability measure, is bounded to one, monotonic and continuous from above and below.
@@
 The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure, and is therefore additive.
-This again solidifies the connection between classical probability, additivity of measure and mutual exclusivity.
+This again solidifies the connection between classical probability, additivity of measures and mutual exclusivity.
 
@@
 From this, we can define a state capacity, which is again a sub-additive measure.
-It reduces to the Liouville measure in the classical case and the dimensionality of the Hilbert space in the quantum case.
+It reduces to the Liouville measure in the classical case and the dimensionality of subspaces in the quantum case.
 
@@
 A final remark before we start.
-Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects.
-Therefore, it is important to us that this work is a piece of a much larger puzzle, which connects cleanly and directly to  other areas, such as information geometry, functional analysis, quantum logic and so on.
-It is the hallmark of a true fundamental structure for physics that such connections can be easily made.
-Unfortunately, the nature of scientific publication requires us to publish a narrow slice of the work to fit the scope of a particular journal.
-Ironically, this means that the most important motivations for our approach are out of scope of this article, as we will concentrate only to the connections to measure theory.
+Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects, such as information geometry, functional analysis, quantum logic and so on.
+However, such a complete work would be effectively overwhelming both to produce and to read.
+The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source/access work.
 

### 2025-08-08 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Revert "NonAdditiveProbability: Label cases in figure"
This reverts commit a3d28bd13108c85a112c50c6e5451dc687911549.

--- 2025-08-05 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 	% dimensions
-	\newcommand{\s}{-0.4}
+	\newcommand{\s}{-0.5}
 	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
@@
 
-		% cut
+		% cut 1
+		\begin{scope}[canvas is zx plane at y=\radius-\s]
+			\filldraw[color=colora, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+		\end{scope}
+
+		% cut 2
+		\begin{scope}[canvas is zx plane at y=-\s]
+			\filldraw[color=colorb, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[color=colorb, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[color=colorb, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+		\end{scope}
+
+		% cut 3
 		\begin{scope}[canvas is zx plane at y=\s]
-			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
-			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+			\filldraw[color=colorc, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[color=colorc, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[color=colorc, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
 		\end{scope}

### 2025-08-08 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Turn equation into figure

--- 2025-08-07 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
-Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$ (see \eqref{eq:dice} for an illustration), where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/5$ of $\ens$ can be constructed from $A$, but no more.
+Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$ (see fig.
+\ref{fig:dice} for an illustration), where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/5$ of $\ens$ can be constructed from $A$, but no more.
 Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
@@
 
-\begin{equation}\label{eq:dice}
+\begin{figure}
 	\def\xstep{0.75em}
@@
 	\end{tikzpicture}
-	= \frac{1}{6}
+	$= \frac{1}{6}$
 	\begin{tikzpicture}[y=3ex, x=\xstep]
@@
 	\end{tikzpicture}
-	+ \frac{5}{6}
+	$+ \frac{5}{6}$
 	\begin{tikzpicture}[y=3ex, x=\xstep]
@@
 	\end{tikzpicture}
-\end{equation}
+	\caption{Illustration of the mixture of distributions}\label{fig:dice}
+\end{figure}
 

### 2025-08-08 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Add old dice example
Since I accidentally changed the previous example instead of adding a new one, I now recreated the initial example again.
NonAdditiveProbability: placeholders for citations

@@
 
-In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts, which stems from the underlying aim to characterize physical theories from an operational viewpoint.
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts, [CITE - GPT, Ariel, Reginatto/Hall...] which stems from the underlying aim to characterize physical theories from an operational viewpoint.
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
-Quasiprobabilistic approaches are also used in quantum mechanics, but it is not completely clear how exactly they relate to standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to standard one.
 
@@
 
-We will first review how measure theory is used in classical probability and classical statistical mechanics.
-We will connect this to Choquet theory, which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.
+We will first review how measure theory is used in classical probability [cite] and classical statistical mechanics[cite].
+We will connect this to Choquet theory[cite phelps], which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.
 each ensemble can be represented by a probability measure over pure states).
@@
 We will then review some results from the use of quasi-probability measures.
-The most used is the Wigner function, that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through normal integral.
-While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.
-The key insight is that it is not negative probability that characterizes quasi-probability,\footnote{Some representation guarantee the quasi-probability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
+The most used is the Wigner function,[cite] that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through normal integral.
+While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.[cite] The key insight is that it is not negative probability that characterizes quasi-probability,\footnote{Some representation guarantee the quasi-probability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
 The quasi-probability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
@@
 
-Lastly, we lay down the requirements for a non-additive generalization of probability theory.
-While should be in principle possible, we show that the most used non-additive integral, the Sugeno and Choquet integral, do not work for our purposes.
+Lastly, we lay down the requirements for a non-additive generalization of probability theory, which should be in principle possible.
+Since there is a wide literature on the subject, [cite] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
 This means that more work will be needed to develop the appropriate tools.
@@
 However, such a complete work would be effectively overwhelming both to produce and to read.
-The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source/access work.
+The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source/access work.[cite our book]
 

### 2025-08-12 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed with Christine

@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to standard one.
-
-The paper aims to ask the following question: can we, by arguing from necessity, find a minimal set of physical requirements for a state space of a physical system, that allows to directly connect to all the measure theoretic structures currently used in physics (i.e.
-probability measures, Liouville measure to count states, quasi-probability, ...) and point to a straight-forward generalization of such structures?
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+
+The paper aims to ask the following question: can we, by arguing from necessity, find a minimal set of physical requirements for a state space of a physical system that allows us to directly connect to all the measure theoretic structures currently used in physics (i.e.
+probability measures, Liouville measure to count states, quasi-probability, ...) and point to a straightforward generalization of such structures?
 We will show that this is indeed possible.\footnote{Note that we believe that the requirements we set, though necessary, are not sufficient to generalize the Lie algebra given by the classical Poisson brackets and the quantum commutators, meaning they cannot fully recover mechanics.
@@
 the collection of all possible preparation procedures) that the theory allows.
-As a basic requirement, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from GPTs is the entropy requirement.
-We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough real valued quantities to identify each ensemble (i.e.
+As a basic requirement, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from Generalized Probabilistic Theories (GPTs) is the entropy requirement.
+We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough experimentally measurable quantities to identify each ensemble (i.e.
 there is a family of affine functions that identify each ensemble through their values).
 Mathematically, these premises lead us to assume that an ensemble space is a convex subset of a metrizable locally convex topological vector space.
-The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the dimension of orthogonal subspaces of an operator)?
+The question is, then, what generalized measure theoretic tool can quantify the number of states (reducing to the Liouville measure in the classical case) or the number of possible outcomes of a measurement (reducing to the dimension of a subspace in the quantum case)?
 What measure theoretic tool can represent each ensemble?
@@
 Uniqueness requires the space to be a Choquet simplex, the generalization of the finite-dimensional simplex, and represents the classical case.
-The key insight, then, is that an ensemble space is classical if and only if each ensembles admits a unique decomposition in terms of pure states.
+The key insight, then, is that an ensemble space is classical if and only if each ensemble admits a unique decomposition in terms of pure states, it is a Choquet simplex.
 
 We will then review some results from the use of quasi-probability measures.
-The most used is the Wigner function,[cite] that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through normal integral.
-While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.[cite] The key insight is that it is not negative probability that characterizes quasi-probability,\footnote{Some representation guarantee the quasi-probability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
+The most used is the Wigner function,[cite] that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through standard integration.
+While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.[cite] The key insight is that it is not negative probability that characterizes quasi-probability,\footnote{Some representations guarantee the quasi-probability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
 The quasi-probability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
@@
 The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
-Given an ensemble $\ens$, we can ask what fraction can be produced by a mixture of a given ensembles.
-This leads to the fraction capacity: a sub-additive set function that, like the probability measure, is bounded to one, monotonic and continuous from above and below.
+Given an ensemble $\ens$, we can ask what fraction can be produced by a mixture of a given set of ensembles.
+This leads to the fraction capacity: a sub-additive set function that, like the probability measure, is bounded by one, monotonic and continuous from above and below.
 Over the extreme points, the fraction capacity represents the supremum of all possible measures that can represent the same ensemble in Choquet theory.
@@
 While in classical statistical mechanics we start with a measure that quantifies the count of states and define the entropy using its logarithm, for the ensemble space we proceed in the opposite way: we start from the entropy and we define the count of states as the exponential of the entropy (i.e.
-the entropic volume).
-From this, we can define a state capacity, which is again a sub-additive measure.
+the entropic volume).[cite Hall] From this, we can define a state capacity, which is again a sub-additive measure.
 It reduces to the Liouville measure in the classical case and the dimensionality of subspaces in the quantum case.
@@
 However, such a complete work would be effectively overwhelming both to produce and to read.
-The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source/access work.[cite our book]
+The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source work.[cite our book]
 
@@
 \textbf{Basic requirements for an ensemble space}.
-It is a cardinal rule in Physical Mathematics that all axioms and mathematical definitions must be grounded in physical requirements.
-This way we can develop mathematical frameworks that are physically well founded.
-As part of our larger project, we are developing a general theory of ensemble spaces, the full treatment of which is beyond the scope of this work.
-We limit ourselves to give the initial premises and the final mathematical characterization, without justifications and proofs.\footnote{More details are available in our ongoing work (cite)}
+In our larger work, we strive to give all axioms and mathematical definitions solid grounding in physical requirements.\footnote{We call this approach Physical Mathematics, as starting from physical ideas recovers the mathematical structures, and only the mathematical structures, that are physically significant.} This way we can develop mathematical frameworks that are physically well founded.
+Specifically, we are developing a general theory of ensemble spaces, the full treatment of which is beyond the scope of this work.
+Here we limit ourselves to giving only the initial premises and the final mathematical characterization, with more details in the appendix and in our ongoing work.[cite book]
 
@@
 A physical law, then, ultimately expresses relationships among ensembles.
-A minimal requirement for a physical theory, then, is to provide an ensemble space: the collection of all possible ensembles within the theory, together with a basic mathematical structure that all physical theories must possess.
+A minimal requirement for a physical theory, then, is to provide an ensemble space: the collection of all possible ensembles within that theory, together with a basic mathematical structure that all physical theories must possess.
 
@@
 First, it needs to guarantee experimental verifiability.
-Mathematically, the ensemble space must be a $\mathsf{T}_0$ second countable topological space, where each open set represents an experimentally verifiable statement.\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Second, it needs to allow statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.
-Mathematically, this constrains the ensemble space to be a convex set.\footnote{Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i}^{\infty} p_i \ens_i$ converges in the space or not.} Third, an entropy function must be defined that characterizes the variability of the elements within an ensemble.
+Mathematically, the ensemble space must be a $\mathsf{T}_0$ second countable topological space, where each open set represents an experimentally verifiable statement[cite book and kevin kelly].\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Second, it needs to allow statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.
+Mathematically, this constrains the ensemble space to be a convex set.\footnote{Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i=1}^{\infty} p_i \ens_i$ converges in the space or not.} Third, an entropy function must be defined that characterizes the variability of the elements within an ensemble.
 
@@
 The last requirement is that there exists a family $F_i$ of statistical quantities that fully identify each ensemble.
-That is, $F_i(p \ens[a] + \bar{p} \ens[b]) = p F_i(\ens[a]) + \bar{p} F_i(\ens[b])$ and $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.\footnote{For example, on the continuum this means that the moment problem is solvable: you can identify the distribution by the expectation of the polynomials of position and momentum.
-Similarly, in quantum mechanics means that the state can be reconstructed through quantum tomography.}
+That is, $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.
 
@@
 Therefore, while the above requirements are needed for a physical theory, it does not necessarily mean that it is convenient to describe problems or make calculations in terms of those definitions.
-In fact, those requirements interact with each other leading to a rich mathematical structure that connects to many of the structures we use in mathematical physics, and its goal is mainly to see how the different representations are connected.
-
-The goal of this paper is to examine the different ways that measure theory can be used to represent ensembles in any given physical theory.
+In fact, those requirements interact with each other leading to a rich mathematical structure that connects to many of the structures we use in mathematical physics, and our goal is mainly to see how the different representations are connected.
+In particular, here we want to examine the different ways that measure theory can be used to represent ensembles in any given physical theory.
 
@@
 
-\textbf{Ensemble spaces as convex sets.} For the purpose of measure theoretic representations, it is best to understand an ensemble space $\Ens$ as a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{The entropy constraint forces the topology to be $\mathsf{T}_1$ the convex space to embed into a vector space.
-A second countable $\mathsf{T}_1$ TVS is metrizable.
-The family of statistical quantities induces a family of semi-norms on the space, which must be countable due to second countability, which makes the TVS locally convex.} The entropy is a function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
+\textbf{Ensemble spaces as convex sets.} For the purpose of measure theoretic representations, it is best to understand an ensemble space $\Ens$ as a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{To see the connection between the requirements and this mathematical characterization see the appendix.} See Fig.~\ref{convex_examples} for examples.
+The entropy is a function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
 \begin{equation}
 	\begin{aligned}
-		p S(\ens[a]) + (1-p) &S(\ens[b]) \leq S(p \ens[a] + (1-p) \ens[b]) \\
-		&\leq I(p,\bar{p}) + p S(\ens[a]) + (1-p) S(\ens[b])
+		p S(\ens[a]) + \bar{p} &S(\ens[b]) \leq S(p \ens[a] + \bar{p} \ens[b]) \\
+		&\leq I(p,\bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])
 	\end{aligned}
 \end{equation}
-where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{In the larger work, we can show that the Shannon entropy is the only upper bound possible.}  Intuitively, the variability (i.e.
+where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{The Shannon entropy is recovered as the only upper bound possible.}  Intuitively, the variability (i.e.
 the entropy) cannot decrease during mixing, and will have an upper bound when mixing two ensembles whose instances are always distinguishable.
@@
 	\end{tikzpicture}
-	\caption{Examples of a classical discrete ensemble space with three pure states (triangle) and of a two dimensional quantum ensemble space (ball).}
+	\caption{Examples of a classical discrete ensemble space with three pure states (triangle) and of a two dimensional quantum ensemble space (ball).}\label{convex_examples}
 \end{figure}
@@
 This will recover the notion of orthogonality for both classical and quantum spaces.
-Conceptually, these correspond to two ensembles where all instances are distinct from the other: they are mutually exclusive and a single instance is enough to distinguish between the two cases.
-
-In classical spaces, orthogonality and separatedness coincide: two ensembles are orthogonal and separate if and only if the support of the respective probability distributions overlap on a non-measure-zero set.
-In quantum mechanics, these two notions are different: any two pure states are separate but are not orthogonal.
-For example, pure ensembles of spin up and spin down are orthogonal, mutually exclusive, precisely because it takes one instance to determine which one is which.
-Pure ensembles of spin up and spin right, however, are not orthogonal precisely because, in some cases, their instance cannot be perfectly discriminated.
-They are, however, orthogonal, because they are both pure states, one is not the mixture of other ensembles: they are extreme points.
+Conceptually, these correspond to two ensembles where all instances are distinct from the other: they are mutually exclusive and a single instance is enough to distinguish between the two ensembles.
+
+In classical spaces, orthogonality and separatedness coincide: two ensembles are orthogonal and separate if and only if the supports of the respective probability distributions overlap on a non-measure-zero set.
+In quantum mechanics, these two notions are different: any two pure states are separate but are not necessarily orthogonal.
+For example, pure ensembles of spin up and spin down are orthogonal (i.e.
+mutually exclusive) precisely because it takes one instance to determine which one is which.
+Pure ensembles of spin up and spin right, however, are not orthogonal precisely because their instances cannot be perfectly discriminated.
+They are, however, separate, because they are both pure states, one is not the mixture of other ensembles: they are extreme points.
+
+TODO: add figure that shows separatedness/orthogonality in classical and quantum.
 
@@
 
-\textbf{Expectation and integration.} A random variable is a function $F : X \to \mathbb{R}$ that assigns a value to each state.\footnote{The function $f$ must be measurable, in the sense that the inverse of a Borel set should be a Borel set.
+\textbf{Expectation and integration.} A random variable is a function $F : X \to \mathbb{R}$ that assigns a value to each state.\footnote{The function $F$ must be measurable, in the sense that the inverse of a Borel set should be a Borel set.
 We find that the term is confusing for a physics audience, so it is avoided in the main text.} The number of the side of the die, position, and momentum are typical examples.
@@
 $E[aF+bG]= aE[F]+bE[G]$).
-In our case, what is more interesting is that it is a convex function in terms of probability measures: if $p = \lambda p_1 + \bar{\lambda} p_2$ then $\int_X F dp = \lambda \int_X F dp_1 + \bar{\lambda} \int_X F dp_2$.
+In our case, what is more interesting is that it is an affine of the probability measure: if $p = \lambda p_1 + \bar{\lambda} p_2$ then $\int_X F dp = \lambda \int_X F dp_1 + \bar{\lambda} \int_X F dp_2$.
 That is, as we required for statistical variables, the expectation for a mixture is the mixture of the expectations.
@@
 This cannot be generalized to quantum mechanics and beyond.
-In our theory of ensemble spaces we start with ensembles, and have to show when and how they can be thought, can be represented, as probability measures.
+In our theory of ensemble spaces we start with ensembles, and have to show when and how they can be thought of, can be represented, as probability measures.
 If we can assume that $E$ is a compact convex set, Choquet theory answers this question.
@@
 In other words, the set of states $X$ is recovered as the set of ensembles with the narrowest spread.
-We say that $\ens \in E$ is represented by a probability measure $p : \Sigma_{E} \to [0,1]$ if $f(\ens) = \int_E f dp$ for every continuous linear functional $f$ on $V$.
+We say that $\ens \in E$ is represented by a probability measure $p : \Sigma_{E} \to [0,1]$ if $F(\ens) = \int_E F dp$ for every continuous linear functional $F$ on $V$.
 Moreover, $p$ is supported by the extreme points if $p(X) = 1$.
-Physically, we start with an ensemble $\ens$ taken as a point in our ensemble space, identified by the value of all possible statistical variables $f(\ens)$, and we represent it as a probability measure $p$, a statistical mixture, over other the extreme points, the states, recovering the expectation with an integral.
+Physically, we start with an ensemble $\ens$ taken as a point in our ensemble space, identified by the value of all possible statistical variables $F(\ens)$, and we represent it as a probability measure $p$, a statistical mixture, over the extreme points, the states, recovering the expectation with an integral.
 
@@
 \begin{thrm}[Choquet 1]
-	Let $E$ be a metrizable compact convex subset of a locally convex space $V$, and let $\ens \in E$.
+	Let $E$ be a metrizable compact convex subset of a locally convex TVS $V$, and let $\ens \in E$.
 Then there exists a probability measure $p_{\ens}$ on $X$ which represents $\ens$ and is supported by the extreme points $X$.
@@
 \begin{thrm}[Choquet 2]
-	Let $E$ be a closed convex metrizable subset of a locally convex space $V$.
+	Let $E$ be a closed convex metrizable subset of a locally convex TVS $V$.
 Then $E$ is a Choquet simplex if and only if for each $\ens \in E$ there exists a unique measure $p_{\ens}$ which represents $\ens$ and is supported by the extreme points $X$.
 \end{thrm}
-The Choquet simplex is a generalization of the finite dimensional simplex, whose mathematical characterization is rather technical.
-The key point is that $E$ is a Choquet simplex if and only if the ensemble space does not allow multiple decomposition in terms of pure states.
+The Choquet simplex, whose mathematical characterization is rather technical, is a generalization of the finite dimensional simplex.
+The key point is that $E$ is a Choquet simplex if and only if the ensemble space does not allow multiple decompositions in terms of pure states.
 This confirms the physical intuition from classical and quantum statistical mechanics.
 
-For example, the ensemble space for a two state quantum system, the Bloch ball, is a compact convex subset.
+For example, the ensemble space for a two state quantum system, the Bloch ball, is a compact convex subset of $\mathbb{R}^3$.
 Any internal point can be seen as the convex combination of points on the surface, the pure states.
-The center, which represents the maximally mixed state, can be thought as an equal mixture of any two opposite points, or a uniform distribution over the whole surface.
-The fact that any ensemble can be seen as a mixture of a set of pure states, then, is reflected by the first part of the Choquet theorem, and it is valid for all physical theories.
+The center, which represents the maximally mixed state, can be thought of as an equal mixture of any two opposite points, or a uniform distribution over the whole surface.
+The fact that any ensemble can be seen as a mixture of a set of pure states, then, is reflected by the first Choquet theorem, and it is valid for all physical theories.
 Uniqueness of the representation fails, however, precisely because the ensemble space is not a simplex: it is not a classical space.
@@
 
-\textbf{Quasi-probability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions, of which the Wigner function is probably the most widely known.
+\textbf{Quasiprobability representations.} Another way to represent quantum ensembles is through the use of quasiprobability distributions, of which the Wigner function is probably the most widely known.
 The Wigner transform $W : \Ens \to L^1(\mathbb{R}^2, \mathbb{R}) $, defined as \cite[eq.~11.8-3a]{mandel1995}
 \begin{equation}
-	W(\rho) = W_{\rho} (q, p) = 2 \int_{-\infty}^{\infty} \langle x + y | \rho | x - y \rangle e^{-2 i p y / \hbar} \; \mathrm dy,
+	W(\rho) = W_{\rho} (x, p) = 2 \int_{-\infty}^{\infty} \langle x + y | \rho | x - y \rangle e^{-2 i p y / \hbar} \; \mathrm dy,
 \end{equation}
@@
 \begin{equation}
-	\tr \left( \frac{1}{2}[XP+PX] \rho \right) = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} xp \; W_{\rho} (q, p) dq dp \; .
-\end{equation}
-Moreover, the transform is an affine map: $W(p \rho_1 + \bar{p} \rho_2 ) = p W(\rho_1) + \bar{p} W(\rho_2)$.
+	\tr \left( \frac{1}{2}[XP+PX] \rho \right) = \int_{-\infty}^{\infty} \int_{-\infty}^{\infty} xp \; W_{\rho} (x, p) dx dp \; .
+\end{equation}
+Moreover, the transform is an affine map: $W(\lambda \rho_1 + \bar{\lambda} \rho_2 ) = \lambda W(\rho_1) + \bar{\lambda} W(\rho_2)$.
 
 In some problems, especially in quantum optics, normal ordering and anti-normal ordering are more useful.
-The Glauber P representation and the Husimi Q representation work like the Wigner function except that the expectations of polynomials recover the respective order.
+The Glauber P [cite] representation and the Husimi Q [cite] representation work like the Wigner function except that the expectations of polynomials recover the respective order.
 In other circumstances, quasi-probability distributions over observables different from position and momentum are more useful.
-For example, the Kirkwood-Dirac is used to represent two-time correlation functions over observables defined at different times \cite{lostaglio2023}.
+For example, the Kirkwood-Dirac quasiprobability is used to represent two-time correlation functions over observables defined at different times \cite{lostaglio2023}.
 Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
-For example, the SIC-POVM represents quantum states as affine combinations of pure states $P_i$ that are ``equally distant'' from each other, which means that no two elements are orthogonal to each other.
-Each ensemble $\rho$ can then be represented as \cite[eqs.
-28-29]{fuchs2004}
-\begin{equation}
-	\rho = \sum_{n=1}^{d^2} \lambda_i P_i = \sum_{n=1}^{d^2} \left[ (d^2 + 1) p_i - \frac{1}{d} \right] P_i \; ,
-\end{equation}
-where $p_i = \tr \left(P_i \rho\right)$.
+For example, the SIC-POVM \cite{fuchs2004} represents quantum states as affine combinations of pure states $P_i$ that are ``equally distant'' from each other, see Fig.
+\ref{SIC}, which means that no two elements are orthogonal to each other.
+Each ensemble $\rho$ can then be represented as
+\begin{equation}
+	\rho = \sum_{i=1}^{d^2} \lambda_i P_i = \sum_{i=1}^{d^2} \left[ (d^2 + 1) p_i - \frac{1}{d} \right] P_i \; ,
+\end{equation}
+where $p_i = \tr \left(P_i \rho\right)$ and $d$ is the dimensionality of the Hilbert space.
 
@@
 The space $Z$, in fact, can change from representation to representation: it is $\mathbb{R}^{2n}$ for a Wigner function, $\mathbb{C}^n$ for the Husimi Q and $[1, d^2] \subset \mathbb{N}$ for the SIC-POVM.
-All representations preserves convex combinations: $W(\sum_i p_i \rho_i) = \sum_i p_i W(\rho_i)$.
+All representations preserve convex combinations: $W(\sum_i \lambda_i \rho_i) = \sum_i \lambda_i W(\rho_i)$.
 That is, a statistical mixture of quasi-probability distributions represents the statistical mixture of the corresponding states.
@@
 States with positive $\lambda_i$ lie within the tetrahedron, while all other states require negative coefficients.
-(b) The Bloch ball can alternatively be inscribed within a larger tetrahedron to guarantee positive coefficients, but the extreme points are no longer physical states.}
+(b) The Bloch ball can alternatively be inscribed within a larger tetrahedron to guarantee positive coefficients, but the extreme points are no longer physical states.}\label{SIC}
 \end{figure}
@@
 That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.\footnote{Since the Husimi Q is defined as $Q_{\rho} ( \alpha ) = \frac{1}{\pi} \langle \alpha | \rho | \alpha \rangle$, it is proportional to the probability of measuring a coherent (i.e.
-Gaussian state) $| \alpha \rangle$, which is non-negative.
+Gaussian) state $| \alpha \rangle$, which is non-negative.
 } Similarly, a SIC-POVM characterizes the Bloch ball by four pure states that form a tetrahedron inscribed in the ball.
@@
 Conversely, we can use signed measures to represent classical distributions.
-Suppose we have a two state classical discrete space, which means the space of all possible ensembles is spanned by $[p, (1-p)]$.
-Alternatively, we can write it as $p[1,0] + (1-p)[0,1]$ which highlights that we are writing each possible probability distribution in terms of the two possible cases.
-But we could also express each probability distribution in terms of other two distributions, as in $\left(2p - \frac{1}{2}\right)\left[\frac{3}{4}, \frac{1}{4}\right] + \left(\frac{3}{2} - 2p\right)\left[\frac{1}{4}, \frac{3}{4}\right]$.
-The coefficients now span between $\left[-\frac{1}{2}, \frac{3}{2}\right]$, can be negative, but still sum to one.
+Suppose we have a two state classical discrete space, which means the space of all possible ensembles is spanned by the vector $\begin{bmatrix} p & (1-p) \end{bmatrix}$.
+Alternatively, we can write it as $p\begin{bmatrix}1 & 0\end{bmatrix} + (1-p)\begin{bmatrix}0 & 1\end{bmatrix}$ which highlights that we are writing each possible probability distribution in terms of the two possible cases.
+But we could also express each probability distribution in terms of two other distributions, as in $\left(2p - \frac{1}{2}\right)\begin{bmatrix}\frac{3}{4} & \frac{1}{4}\end{bmatrix} + \left(\frac{3}{2} - 2p\right)\begin{bmatrix}\frac{1}{4} & \frac{3}{4}\end{bmatrix}$.
+TODO: check.
+The coefficients now span between $\begin{bmatrix}-\frac{1}{2}, \frac{3}{2}\end{bmatrix}$, can be negative, but still sum to one.
 That is, in terms of this new affine basis, we need to use a signed probability.
@@
 \end{equation}
-where $F_\psi$ is an entire analytic function, which cannot vanish on an open set without vanishing everywhere, therefore the support of all Husimi Q functions is the whole space $Z$.
-This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support; moreover disjoint subsets of $Z$ cannot be thought of as mutually exclusive events because we are not going to be able to find two distributions with those subsets as their respective supports.
+where $F_\psi$ is an entire analytic function, which cannot vanish on an open set without vanishing everywhere, therefore the support of all Husimi Q functions is the whole space $Z_Q$.
+This means that orthogonal states, which correspond to mutually exclusive events in quantum theory, always have the same support over the space $Z_Q$ that supports Husimi Q functions; moreover disjoint subsets of $Z_Q$ cannot be thought of as mutually exclusive events because we are not going to be able to find two distributions with those subsets as their respective supports.
 For the discrete case, a SIC-POVM representation picks $d^2$ pure states $P_i$ in a $d$ dimensional space, which means they cannot all be orthogonal to each other.
@@
 Therefore, disjoint sets of $P_i$ do not correspond to mutually exclusive events.
-Quasi-probability representations, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
+Quasiprobability representations, then, fail to be true probabilities not because of negative values, but because the points of the space do not represent mutually exclusive events.
+
+TODO: remove all the - in quasi-probability
 
@@
 
-In Choquet theory, we saw that classical spaces, Choquet simplexes, are those in which there is a one-to-one correspondence between probability measures and ensembles, where no multiple decomposition in terms of pure states is allowed.
+In Choquet theory, we saw that classical spaces, Choquet simplexes, are those in which there is a one-to-one correspondence between probability measures and ensembles, where multiple decomposition in terms of pure states is not allowed.
 We want to show that this is a consequence of assuming mutual exclusivity between the points of $X$.
@@
 For the Bloch ball, a classical context is an axis that connects two opposite points.
-The points are orthogonal because the midpoint is the maximally mixed state, which therefore maximizes the entropy increase, and the line segment is a one simplex.
-If we take the position operator and consider all density operators that commute with $x$, this is corresponds to the set of functions $\rho(x)$ that integrate to one, which is the space of probability supported by $X$, which is a simplex.
+The points are orthogonal because the midpoint is the maximally mixed state, which therefore maximizes the entropy increase, and the line segment is a 1-simplex.
+If we take the position operator and consider all density operators that commute with $x$, this corresponds to the set of functions $\rho(x)$ that integrate to one, which is the space of probability supported by $X$, which is a simplex.
 
@@
 Now let $M_1^{\pm}(Z)$ be the space of signed probability measures.
-While it is not a subspace of $M(Z)$, it does not include the zero measure, it is an affine subspace therefore, up to the choice of an arbitrary origin, it is a vector space.
+While it is not a subspace of $M^{\pm}(Z)$ as it does not include the zero measure, it is an affine subspace.
+Therefore, up to the choice of an arbitrary origin, it is a vector space.
 
 Recall that an ensemble space must be a convex subset of a vector space.
-If we are able to represent ensembles as quasi-probability distributions over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such an $Z$ always exists is still open for us.
+If we are able to represent ensembles as quasi-probability distributions over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such a $Z$ always exists is still open for us.
 A possible conjecture is that $Z$ could be constructed from the statistical quantities $F_i$ that fully identify each ensemble in $\Ens$.} In this case, any element $v\in V$ can be represented uniquely by a signed probability measure $\mu_{v}$, regardless of whether it is an actual ensemble or not.
-Since statistical quantities $F_i$ are linear functional, we will have $F_i(v) = \int_Z F_i(z) d\mu_{v}$.
+Since statistical quantities $F_i$ are affine functions, we will have $F_i(v) = \int_Z F_i(z) d\mu_{v}$.
 In other words, there is nothing special about quantum mechanics that allows quasi-probability representations.
@@
 
-\textbf{Non-uniqueness of quasi-probability.} Our approach gives us a consistent view of both quasi-probabilities and classical probability.
+\textbf{Non-uniqueness of quasiprobability.} Our approach gives us a consistent view of both quasiprobabilities and classical probability.
 The first are affine representations of the ensembles that connect to the vector space structure, while the second are convex representations of the extreme points which rely on the shape of the ensemble space.
 
-Quasi-probability distributions are very useful in practice, particularly because one can construct them to fit particular problems.
+Quasiprobability distributions are very useful in practice, particularly because one can construct them to fit particular problems.
 Once one has identified a set of observables that are the most natural ones to define a problem, one may find a representation that is specifically suited for the case.
@@
 
+We have seen that classical probability arises exactly when we are describing a set of ensembles that can always be decomposed in a unique set of mutually distinguishable cases.
+In quantum mechanics, instead, pure states
+
 \section*{Acknowledgments}

### 2025-08-13 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: worked on some TODOs

@@
 The paper aims to ask the following question: can we, by arguing from necessity, find a minimal set of physical requirements for a state space of a physical system that allows us to directly connect to all the measure theoretic structures currently used in physics (i.e.
-probability measures, Liouville measure to count states, quasi-probability, ...) and point to a straightforward generalization of such structures?
+probability measures, Liouville measure to count states, quasiprobability, ...) and point to a straightforward generalization of such structures?
 We will show that this is indeed possible.\footnote{Note that we believe that the requirements we set, though necessary, are not sufficient to generalize the Lie algebra given by the classical Poisson brackets and the quantum commutators, meaning they cannot fully recover mechanics.
@@
 
-We will then review some results from the use of quasi-probability measures.
+We will then review some results from the use of quasiprobability measures.
 The most used is the Wigner function,[cite] that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through standard integration.
-While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.[cite] The key insight is that it is not negative probability that characterizes quasi-probability,\footnote{Some representations guarantee the quasi-probability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
-The quasi-probability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
+While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.[cite] The key insight is that it is not negative probability that characterizes quasiprobability,\footnote{Some representations guarantee the quasiprobability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
+The quasiprobability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
 Therefore, while signed measures are a powerful tool for calculation in specific problems, they cannot give us a general principled representation of the physical theory.
@@
 Classical mechanics, for example, is typically described in terms of the geometry of phase space.
-Quantum mechanics is most often defined in terms of complex inner product spaces, but can be reframed in terms of quasi-probability distributions.
+Quantum mechanics is most often defined in terms of complex inner product spaces, but can be reframed in terms of quasiprobability distributions.
 Therefore, while the above requirements are needed for a physical theory, it does not necessarily mean that it is convenient to describe problems or make calculations in terms of those definitions.
@@
 
-%Goal of this section: quasi-probability representations are allowed by the embedding in a topological vector space.
+%Goal of this section: quasiprobability representations are allowed by the embedding in a topological vector space.
 Choice of representations are guided by what observables are interesting for a specific problem.
@@
 
-%Thing to say: all these representations are useful because they preserve the linear structure, meaning that convex combinations of ensembles are convex combinations of quasi-probability; expectation values are linear.
+%Thing to say: all these representations are useful because they preserve the linear structure, meaning that convex combinations of ensembles are convex combinations of quasiprobability; expectation values are linear.
 Check, but it should be true.
 
-% 1 describe quasiprobability approaches; go through examples; point out that the quasi-probability can be negative; that one can recover expectations;
+% 1 describe quasiprobability approaches; go through examples; point out that the quasiprobability can be negative; that one can recover expectations;
 
@@
 Recall that an affine combination is a linear combination where all coefficients sum to one but are not necessarily non-negative.
-Therefore, in general, quasi-probability distributions can be understood as affine combinations.
+Therefore, in general, quasiprobability distributions can be understood as affine combinations.
 Remarkably, expectation values of polynomials of position and momentum in symmetric ordering coincide with expectations over the Wigner function.
@@
 The Glauber P [cite] representation and the Husimi Q [cite] representation work like the Wigner function except that the expectations of polynomials recover the respective order.
-In other circumstances, quasi-probability distributions over observables different from position and momentum are more useful.
+In other circumstances, quasiprobability distributions over observables different from position and momentum are more useful.
 For example, the Kirkwood-Dirac quasiprobability is used to represent two-time correlation functions over observables defined at different times \cite{lostaglio2023}.
-Quasi-probability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
+Quasiprobability representations are also used over discrete elements that may not even correspond to eigenstates of a particular operator.
 For example, the SIC-POVM \cite{fuchs2004} represents quantum states as affine combinations of pure states $P_i$ that are ``equally distant'' from each other, see Fig.
@@
 
-It should be clear that there are many ways to represent quantum states via quasi-probability distributions, and that the choice of representation is a matter of convenience for the problem at hand.
+It should be clear that there are many ways to represent quantum states via quasiprobability distributions, and that the choice of representation is a matter of convenience for the problem at hand.
 All representations, though, have the following features in common.
-First, we have a map $W : \Ens \to Q(Z,\mathbb{R})$ that given each mixed state $\rho \in \Ens$ returns the corresponding quasi-probability function $w(z) \in Q(Z,\mathbb{R})$, where $Q(Z,\mathbb{R}) \subset L^1(Z,\mathbb{R})$ is the function space used by the specific representation.
+First, we have a map $W : \Ens \to Q(Z,\mathbb{R})$ that given each mixed state $\rho \in \Ens$ returns the corresponding quasiprobability function $w(z) \in Q(Z,\mathbb{R})$, where $Q(Z,\mathbb{R}) \subset L^1(Z,\mathbb{R})$ is the function space used by the specific representation.
 The space $Z$, in fact, can change from representation to representation: it is $\mathbb{R}^{2n}$ for a Wigner function, $\mathbb{C}^n$ for the Husimi Q and $[1, d^2] \subset \mathbb{N}$ for the SIC-POVM.
 All representations preserve convex combinations: $W(\sum_i \lambda_i \rho_i) = \sum_i \lambda_i W(\rho_i)$.
-That is, a statistical mixture of quasi-probability distributions represents the statistical mixture of the corresponding states.
-Since the map is linear, and expectation values are linear, one can express expectation values as integrals or sums over the quasi-probability $w(z)$.
+That is, a statistical mixture of quasiprobability distributions represents the statistical mixture of the corresponding states.
+Since the map is linear, and expectation values are linear, one can express expectation values as integrals or sums over the quasiprobability $w(z)$.
 
@@
 
-\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasi-probability \cite{adhikary2020}.
+\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasiprobability \cite{adhikary2020}.
 Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
 This is a mischaracterization.
-For example, the Husimi Q representation provides us with a non-negative quasi-probability function.
+For example, the Husimi Q representation provides us with a non-negative quasiprobability function.
 That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.\footnote{Since the Husimi Q is defined as $Q_{\rho} ( \alpha ) = \frac{1}{\pi} \langle \alpha | \rho | \alpha \rangle$, it is proportional to the probability of measuring a coherent (i.e.
@@
 } Similarly, a SIC-POVM characterizes the Bloch ball by four pure states that form a tetrahedron inscribed in the ball.
-Since we can also inscribe the Bloch ball inside a tetrahedron, we can represent all states of a two-state quantum system by a convex combination of the vertices, leading to a non-negative quasi-probability distribution over four elements.
+Since we can also inscribe the Bloch ball inside a tetrahedron, we can represent all states of a two-state quantum system by a convex combination of the vertices, leading to a non-negative quasiprobability distribution over four elements.
 
@@
 But we could also express each probability distribution in terms of two other distributions, as in $\left(2p - \frac{1}{2}\right)\begin{bmatrix}\frac{3}{4} & \frac{1}{4}\end{bmatrix} + \left(\frac{3}{2} - 2p\right)\begin{bmatrix}\frac{1}{4} & \frac{3}{4}\end{bmatrix}$.
-TODO: check.
 The coefficients now span between $\begin{bmatrix}-\frac{1}{2}, \frac{3}{2}\end{bmatrix}$, can be negative, but still sum to one.
@@
 
-Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measures, the difference between classical probability and quantum quasi-probability does not lie in the use of negative probability.
+Since we can represent classical states with signed probability measures, and quantum states with non-negative probability measures, the difference between classical probability and quantum quasiprobability does not lie in the use of negative probability.
 Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events, and all mutually exclusive events are represented by disjoint regions.
 Mutually exclusive events can be prepared and observed independently from the others (i.e.
-they are orthogonal in the sense that they maximize entropy increase during mixture); we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasi-probability in quantum mechanics.
+they are orthogonal in the sense that they maximize entropy increase during mixture); we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasiprobability in quantum mechanics.
 
@@
 
-TODO: remove all the - in quasi-probability
-
 % 3 These representations are possible because of the vector space structure of the ensemble space, and therefore something similar will be possible in ANY physical theory
@@
 But how exactly does mutual exclusivity give us classical probability?
-Moreover, are quasi-probability representations specific to quantum mechanics, or are they more general?
+Moreover, are quasiprobability representations specific to quantum mechanics, or are they more general?
 This is where the perspective from the ensemble space framework allows us to answer these questions directly.
@@
 
-Now the second question: why does quantum mechanics allow quasi-probability representations?
-Let $Z$ be the space over which we are defining our quasi-probability distributions.
+Now the second question: why does quantum mechanics allow quasiprobability representations?
+Let $Z$ be the space over which we are defining our quasiprobability distributions.
 Note that, in general, $Z$ is not the space of states.
@@
 Recall that an ensemble space must be a convex subset of a vector space.
-If we are able to represent ensembles as quasi-probability distributions over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such a $Z$ always exists is still open for us.
+If we are able to represent ensembles as quasiprobability distributions over $Z$, then, $M_1^{\pm}(Z)$ will be isomorphic to the embedding vector space $V$ as topological vector spaces.\footnote{The question of whether such a $Z$ always exists is still open for us.
 A possible conjecture is that $Z$ could be constructed from the statistical quantities $F_i$ that fully identify each ensemble in $\Ens$.} In this case, any element $v\in V$ can be represented uniquely by a signed probability measure $\mu_{v}$, regardless of whether it is an actual ensemble or not.
 Since statistical quantities $F_i$ are affine functions, we will have $F_i(v) = \int_Z F_i(z) d\mu_{v}$.
-In other words, there is nothing special about quantum mechanics that allows quasi-probability representations.
+In other words, there is nothing special about quantum mechanics that allows quasiprobability representations.
 It is the fact that the ensemble space embeds into a vector space, a general fact for all theories, that allows such representations.
@@
 Once one has identified a set of observables that are the most natural ones to define a problem, one may find a representation that is specifically suited for the case.
-While the quasi-probability cannot be interpreted as probability, it still provides a useful tool to describe and understand the problem and give some intuition.
+While the quasiprobability cannot be interpreted as probability, it still provides a useful tool to describe and understand the problem and give some intuition.
 
 However, this practical convenience is also a drawback for finding a more general approach to probability in spaces beyond classical mechanics.
-Crucially, the space $Z$ over which the quasi-probability measures are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tell us which quasi-probability measures represent valid ensembles and which do not.
+Crucially, the space $Z$ over which the quasiprobability measures are defined is not unique, and it is not defined by the structure of the ensemble space and, as a consequence, cannot tell us which quasiprobability measures represent valid ensembles and which do not.
 
 As a comparison, in the previous case all probability measures were given over a uniquely defined set of points, the extreme points of the ensemble space, but the probability measure may not be unique.
-In this case, the quasi-probability associated to each ensemble is unique, but the space over which the quasi-probability is defined is not.
+In this case, the quasiprobability associated to each ensemble is unique, but the space over which the quasiprobability is defined is not.
 
@@
 \textbf{Uniqueness of the fraction capacity.} From a conceptual standpoint, the fraction capacity has a major advantage: it has a clear physical meaning and it is unique.
-Like in Choquet theory and unlike for quasi-probability, the domain over which the measure is defined is fully determined by the system, by the ensemble space.
-On the other hand, unlike Choquet theory and like quasi-probability, once the space is fixed, there is one unique fraction capacity for each ensemble.
+Like in Choquet theory and unlike for quasiprobability, the domain over which the measure is defined is fully determined by the system, by the ensemble space.
+On the other hand, unlike Choquet theory and like quasiprobability, once the space is fixed, there is one unique fraction capacity for each ensemble.
 Moreover, the setup is more general than Choquet theory (i.e.

### 2025-08-15 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Align dice example with book
Instead of e_123456 = e_1 + e_23456 we split into e_3 + e_12456 like we do in the book.

--- 2025-08-08 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts, [CITE - GPT, Ariel, Reginatto/Hall...] which stems from the underlying aim to characterize physical theories from an operational viewpoint.
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts, which stems from the underlying aim to characterize physical theories from an operational viewpoint.
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.
+Quasiprobabilistic approaches are also used in quantum mechanics, but it is not completely clear how exactly they relate to standard one.
 
@@
 The starting point is that any physical law deals with experimentally reproducible relationships.
-This implies a physical theory must give us the space of statistical ensembles (i.e.
+This implies a physical theory must gives us the space of statistical ensembles (i.e.
 the collection of all possible preparation procedures) that the theory allows.
-As a basic requirement, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from GPTs is the entropy requirement.
-We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough real valued quantities to identify each ensemble (i.e.
-there is a family of affine functions that identify each ensemble through their values).
-Mathematically, these premises lead us to assume that an ensemble space is a convex subset of a metrizable locally convex topological vector space.
+As a basic requirements, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from GPTs is the entropy requirement.
+We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough real valued quantities to identify each ensemble (leading to the continuous embedding into a Hausdorff locally convex topological vector space).
+Mathematically, under these premises, an ensembles space is a convex subset of a locally convex topological vector space.
 The question is, then, what measure theoretic tool can quantify the number of states (as in the classical Liouville measure) or the number of possible outcomes of a measurements (as in the dimension of orthogonal subspaces of an operator)?
@@
 
-We will first review how measure theory is used in classical probability [cite] and classical statistical mechanics[cite].
-We will connect this to Choquet theory[cite phelps], which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.
-each ensemble can be represented by a probability measure over pure states).
+We will first review how measure theory is used in classical probability together with some results from Choquet theory.
+This will tell us that elements of a compact convex set can always be represented by a probability measure over its extreme points, the pure states.
 While such representation is always possible, at least in the compact case, it is not unique.
-Uniqueness requires the space to be a Choquet simplex, the generalization of the finite-dimensional simplex, and represents the classical case.
-The key insight, then, is that an ensemble space is classical if and only if each ensembles admits a unique decomposition in terms of pure states.
+Uniqueness requires the space to be a Choquet simplex, which is a generalization of the standard finite-dimensional simplex and represents the classical case.
+The key insight, then, an ensemble space is classical if and only if each ensembles admits a unique decomposition in terms of pure states.
 
 We will then review some results from the use of quasi-probability measures.
-The most used is the Wigner function,[cite] that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through normal integral.
-While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.[cite] The key insight is that it is not negative probability that characterizes quasi-probability,\footnote{Some representation guarantee the quasi-probability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
+While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.
+The key insight is that it is not negative probability that characterizes quasi-probability, but the lack of mutual exclusivity of the points of the space.
 The quasi-probability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
@@
 The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
-Given an ensemble $\ens$, we can ask what fraction can be produced by a mixture of a given ensembles.
+Given an ensemble $\ens$, we can ask what fraction can be produces by a mixture of a given ensembles.
 This leads to the fraction capacity: a sub-additive set function that, like the probability measure, is bounded to one, monotonic and continuous from above and below.
@@
 The classical case exactly corresponds to the simplex case, in which fraction capacity corresponds to a unique measure, and is therefore additive.
-This again solidifies the connection between classical probability, additivity of measures and mutual exclusivity.
+This again solidifies the connection between classical probability, additivity of measure and mutual exclusivity.
 
@@
 From this, we can define a state capacity, which is again a sub-additive measure.
-It reduces to the Liouville measure in the classical case and the dimensionality of subspaces in the quantum case.
-
-Lastly, we lay down the requirements for a non-additive generalization of probability theory, which should be in principle possible.
-Since there is a wide literature on the subject, [cite] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
+It reduces to the Liouville measure in the classical case and the dimensionality of the Hilbert space in the quantum case.
+
+Lastly, we lay down the requirements for a non-additive generalization of probability theory.
+While should be in principle possible, we show that the most used non-additive integral, the Sugeno and Choquet integral, do not work for our purposes.
 This means that more work will be needed to develop the appropriate tools.
@@
 A final remark before we start.
-Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects, such as information geometry, functional analysis, quantum logic and so on.
-However, such a complete work would be effectively overwhelming both to produce and to read.
-The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source/access work.[cite our book]
+Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects.
+Therefore, it is important to us that this work is a piece of a much larger puzzle, which connects cleanly and directly to  other areas, such as information geometry, functional analysis, quantum logic and so on.
+It is the hallmark of a true fundamental structure for physics that such connections can be easily made.
+Unfortunately, the nature of scientific publication requires us to publish a narrow slice of the work to fit the scope of a particular journal.
+Ironically, this means that the most important motivations for our approach are out of scope of this article, as we will concentrate only to the connections to measure theory.
 
@@
 Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
-Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$, where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/6$ of $\ens$ can be constructed from $A$, but no more.
+Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$ (see fig.
+\ref{fig:dice} for an illustration), where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/5$ of $\ens$ can be constructed from $A$, but no more.
 Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
-Note that, given the uniform distribution, $1/6$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
+Note that, given the uniform distribution, $1/5$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
 This gives the basic insight for our definitions.
+
+\begin{figure}
+	\def\xstep{0.75em}
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
+	\end{tikzpicture}
+	$= \frac{1}{6}$
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(3, 1)};
+	\end{tikzpicture}
+	$+ \frac{5}{6}$
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/4) (4, 1/5) (5, 1/5) (6, 1/5)};
+	\end{tikzpicture}
+
+	\vspace{0.3cm}
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
+	\end{tikzpicture}
+	$= \frac{1}{3}$
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(3, 1/2) (4, 1/2)};
+	\end{tikzpicture}
+	$+ \frac{2}{3}$
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/4) (2, 1/4) (5, 1/4) (6, 1/4)};
+	\end{tikzpicture}
+	\caption{Illustration of the mixture of distributions}\label{fig:dice}
+\end{figure}
 
@@
 	% dimensions
-	\newcommand{\s}{-0.5}
+	\newcommand{\s}{-0.4}
 	\pgfmathsetmacro{\radiusats}{sqrt(1 - \s*\s)}
@@
 	\newcommand{\margin}{1.3}
-
-	\definecolor{colora}{HTML}{1EBCB4}
-	\definecolor{colorb}{HTML}{E57C04}
-	\definecolor{colorc}{HTML}{7A5CE3}
 
@@
 
-		% cut 1
-		\begin{scope}[canvas is zx plane at y=\radius-\s]
-			\filldraw[color=colora, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
-		\end{scope}
-
-		% cut 2
-		\begin{scope}[canvas is zx plane at y=-\s]
-			\filldraw[color=colorb, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
-			\draw[color=colorb, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-			\draw[color=colorb, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
-		\end{scope}
-
-		% cut 3
+		% cut
 		\begin{scope}[canvas is zx plane at y=\s]
-			\filldraw[color=colorc, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
-			\draw[color=colorc, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-			\draw[color=colorc, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+			\filldraw[color=blue, thick, draw opacity=1, fill opacity=0.1] (-\margin, -\margin) -- (\margin, -\margin) -- (\margin, \margin) -- (-\margin, \margin) -- cycle;
+			\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+			\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
 		\end{scope}
@@
 		% empty case
-		\draw[color=colora] (3.75*\xx+\xy, 3.75*\yx+\yy, 3.75*\zx+\zy) node {$s > b$};
+		\draw (3.75*\xx+\xy, 3.75*\yx+\yy, 3.75*\zx+\zy) node {$s > b$};
 
@@
 		\begin{scope}[shift={(4.5*\xx-0.2*\xy, 4.5*\yx-0.2*\yy, 4.5*\zx-0.2*\zy)}, scale=0.5]
-			\draw[color=colorb] (-3.5, +1.3, 0) node {$\frac{1}{2} ( a + b ) < s < b$};
+			\draw (-3.5, +1.3, 0) node {$\frac{1}{2} ( a + b ) < s < b$};
 
@@
 			\begin{scope}[canvas is zx plane at y=-\s]
-				\draw[color=colorb, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-				\draw[color=colorb, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
 			\end{scope}
@@
 		\begin{scope}[shift={(3*\xx-\xy, 3*\yx-\yy, 3*\zx-\zy)}, scale=0.5]
-			\draw[color=colorc] (3.5, -0.5, 0) node {$s < \frac{1}{2} ( a + b )$};
+			\draw (3.5, -0.5, 0) node {$s < \frac{1}{2} ( a + b )$};
 
@@
 			\begin{scope}[canvas is zx plane at y=\s]
-				\draw[color=colorc, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
-				\draw[color=colorc, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[blue, thick, dashed] ({\iyz * \radiusats}, {\iyx * \radiusats}) arc[start angle={atan2(\iyx, \iyz)}, delta angle=180, radius=\radiusats cm];
+				\draw[blue, thick] ({-\iyz * \radiusats}, {-\iyx * \radiusats}) arc[start angle={atan2(-\iyx, -\iyz)}, delta angle=180, radius=\radiusats cm];
 			\end{scope}

### 2025-08-19 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed with Christine

--- 2025-08-13 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 Let $\ens_{123456}$ be the uniform distribution over all outcomes.
-Let $A_1 = \{\ens_1\}$ where $\ens_1$ represents outcome one with 100\% probability.
-Since we can write $\ens_{123456} = \frac{1}{6} \ens_{1} + \frac{5}{6} \ens_{23456}$, where $\ens_{23456}$ represents a uniform distributions over the five outcomes, $1/6$ of $\ens$ can be constructed from $A$, but no more.
-Similarly, if $A_{12} = \{\ens_{1},\ens_{2}\}$, we can write $\ens = \frac{1}{3} \left(\frac{1}{2} \ens_1 + \frac{1}{2} \ens_2 \right)  + \frac{2}{3} \ens_{3456}$, so $1/3$ of $\ens$ can be constructed from $A$, but no more.
-Note that, given the uniform distribution, $1/6$ is exactly the probability for the event $A_1$ and $1/3$ the probability for event $A_{12}$.
-This gives the basic insight for our definitions.
+Let $A_3 = \{\ens_3\}$ where $\ens_3$ represents outcome three with 100\% probability.
+Since we can write $\ens_{123456} = \frac{1}{6} \ens_{3} + \frac{5}{6} \ens_{12456}$, where $\ens_{12456}$ represents a uniform distribution over the five other outcomes, $1/6$ of $\ens_{123456}$ can be constructed from $A_3$, but no more.
+Similarly, if $A_{34} = \{\ens_{3},\ens_{4}\}$, we can write $\ens_{123456} = \frac{1}{3} \left(\frac{1}{2} \ens_3 + \frac{1}{2} \ens_4 \right)  + \frac{2}{3} \ens_{1256}$, so $1/3$ of $\ens_{123456}$ can be constructed from $A_{34}$, but no more.
+Note that, given the uniform distribution, $1/6$ is exactly the probability for the event $A_3$ and $1/3$ the probability for event $A_{34}$.
+This gives the basic intuition for our definitions.
 
@@
 
-We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a Borel set of ensembles $A \subset \Ens$.
+We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a Borel set of ensembles $A \subseteq \Ens$.
 Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.
-mixtures),\footnote{Since we work with convex sets within a topology, the hull is both closed under the topology and convex combinations.
-That is, $\hull(A)$ is the smallest closed and convex set that contains $A$} we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
+mixtures), we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
 \begin{equation}
@@
 
-\textbf{Compare to Choquet theory.} In case $E$ is compact, we can restrict the fraction capacity to the set $X$ of extreme points.
-Given a set $A \subset X$, of extreme points, we can find the set $N \subset E$ of all ensembles that can be represented with a probability measure with support $A$.
+\textbf{Comparison to Choquet theory.} In case the ensemble space $E$ is compact, we can restrict the fraction capacity to the set $X$ of extreme points.
+Given a set $A \subseteq X$ of extreme points, we can find the set $M \subseteq E$ of all ensembles that can be represented with a probability measure with support $A$.
 This forms a convex set.
-Now, the ability to express an ensemble $\ens$ as the convex combination of $N$ is precisely the ability to find a measure $p$ that represents $\ens$ such that $p(A) \neq 0$.
-That is, $\ens = \lambda \ens[a] + \bar{\lambda} \ens[b]$ with $\ens[a] \in M$ if and only if there is a probability measure $p$ that represents $\ens$ such that $p(A) \geq \lambda$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{p(A) \, | \, p \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all possible measures.
+Now, the ability to express an ensemble $\ens$ as the convex combination of $M$ is precisely the ability to find a measure $p$ that represents $\ens$ such that $p(A) \neq 0$.
+That is, $\ens = \lambda \ens[a] + \bar{\lambda} \ens[b]$ with $\ens[a] \in M$ if and only if there is a probability measure $p$ that represents $\ens$ such that $p(A) = \lambda$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{p(A) \, | \, p \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all measures that represent $\ens$.
 
@@
 In that case, the fraction capacity restricted to the extreme points will also be additive.
-In the non-classical case, the closure of the ensemble space will not be a simplex, and there will be multiple representations, which will lead to a non-additive measure.
+In the non-classical case, the closure of the ensemble space will not be a simplex, and ensembles will be represented by multiple measures, which will lead to a non-additive fraction capacity.
 Suppose, in fact, that $p_1$ and $p_2$ are two distinct measures that represent the same ensemble $\ens$.
 Then there will be a set $A$ such that $p_1(A) \neq p_2(A)$.
-Suppose, without loss of generality, that $p_1(A) > p_2(A)$ We will also have $p_1(A^{\complement}) = 1- p_1(A) < 1 - p_2(A) = p_2(A^{\complement})$.
-Since the fraction capacity for a set cannot be less than the measure on that set, we have $\frcap_{\ens}(A \cup A^{\complement})=\frcap_{\ens}(X) = 1 = p_1(A) + p_1(A^{\complement}) < (A) + p_2(A^{\complement}) \leq \frcap_{\ens}(A) + \frcap_{\ens}(A^{\complement})$.
+Suppose, without loss of generality, that $p_1(A) > p_2(A)$.
+We will also have $p_1(A^{\complement}) = 1- p_1(A) < 1 - p_2(A) = p_2(A^{\complement})$.
+Since the fraction capacity for a set cannot be less than the measure on that set, we have $\frcap_{\ens}(A \cup A^{\complement})=\frcap_{\ens}(X) = 1 = p_1(A) + p_1(A^{\complement}) < p_1(A) + p_2(A^{\complement}) \leq \frcap_{\ens}(A) + \frcap_{\ens}(A^{\complement})$.
 That is, the fraction capacity is additive over disjoint sets of extreme points in and only in the classical case.
@@
 the ensemble space need not be compact) and stems from necessary physical requirements.
-An in depth study on fraction capacities, then, is guaranteed to be applicable to all physical theories, including those yet to be formulated.
+An in-depth study on fraction capacities, then, is guaranteed to be applicable to all physical theories, including those yet to be formulated.
 
@@
 Given that in statistical mechanics the entropy is the logarithm of the count of states, we expect the count of states to be the exponential of the entropy.
-Therefore one can argue that the proper way to measure the spread of an ensemble is through the exponential of the entropy, also known as entropic volume.\footnote{It has already been shown in \cite{hall1999} that the exponential of the entropy is the only function that can be used to define a volume.} However, the count of states, like the probability measure, is typically defined on the sample space $\Omega$, which can only work in the classical case.
+Therefore one can argue that the proper way to measure the spread of an ensemble is through the exponential of the entropy, also known as entropic volume.\footnote{It has already been shown in \cite{hall1999} that the exponential of the entropy is the only function that can be used to define a volume.} However, the count of states, like the probability measure, is typically defined on the sample space $X$, which can only work in the classical case.
 The entropy volume is a function of the ensemble space, so it is not a measure.
@@
 To build an intuition, consider the classical case.
-For every region $U$ of phase space, we have a set of possible probability distributions that have support $U$.
-This is a convex set since every mixture of distributions with support $U$ has support $U$.
-Note that the uniform distribution $\rho_U$, if it exists, is the one with the greatest entropy, which will correspond to $S(\rho_U) = \log \mu(U)$.
-If $U$ is not compact, then there is no upper limit to the entropy and $\mu(U) = +\infty$.
-Now, let us transfer this intuition on the ensemble space $\ens$.
-If we take a set $A$, we can ask: what is the highest entropy reachable by a mixture of elements of $A$?
-If $A$ happens to be the set of ensembles with support $U$, then the supremum of the entropy is $\log \mu(U)$, which also works in the unbounded case.
+For every region $A$ of phase space, we have a set of possible probability distributions that have support $A$.
+This is a convex set since every mixture of distributions with support $A$ has support $A$.
+Note that the uniform distribution $\rho_A$, if it exists, is the one with the greatest entropy, which will correspond to $S(\rho_A) = \log \mu(A)$.
+If $A$ is not bounded, then there is no upper limit to the entropy and $\mu(A) = +\infty$.
+Now, let us transfer this intuition to the ensemble space $\Ens$.
+If we take a set $M\subseteq \Ens$, we can ask: what is the highest entropy reachable by a mixture of elements of $M$?
+If $M$ happens to be the set of ensembles with support $A$, then the supremum of the entropy is $\log \mu(A)$, which also works in the unbounded case.
 
@@
 In the general case, we will start with the ensemble space $\Ens$ and the entropy $S$.
-For ensembles, we can define the entropic volume\footnote{CITE Hall} to be the exponential of the entropy $2^{S(\ens)}$.
+For ensembles, we can define the entropic volume to be the exponential of the entropy $2^{S(\ens)}$.
 For the entropic volume to quantify the spread of the ensemble, we must have that the spread can at most sum.
 Remarkably, the bounds on the entropy are such that the exponential of the entropy is subadditive.
-That is, $2^{S(p\ens[a]+\bar{p}\ens[b])} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])}$ for any ensemble space.\footnote{It is an open question whether the exponential is the only monotonic function of the entropy that is subadditive.} For the equality to hold, the ensemble must be orthogonal, meaning that two states are distinguishable if and only if they are mutually exclusive.
+That is, $2^{S(p\ens[a]+\bar{p}\ens[b])} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])}$ for any ensemble space.\footnote{It is an open question whether the exponential is the only monotonic function of the entropy that is subadditive.} For the equality to hold, the ensembles must be orthogonal, meaning that states are distinguishable if and only if they are mutually exclusive.
 Therefore, this is indeed a general way to define the spread of ensembles and the count of distinguishable states in a way that is consistent with the previous discussions.
@@
 
-One can then show that the fraction capacity $\capacity : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
+One can then show that the state capacity $\capacity : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
 \begin{enumerate}
@@
 In the discrete case, every point is associated with zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
-In the continuous case, gives a subset of phase space $U$ and $A_U$ the set of distributions with support within $U$, the maximum entropy is reached by the uniform distribution $\rho_U \in A_U$, which means we recover $\log_2 \capacity(A_U) = S(\rho_U)$, the fundamental postulate of statistical mechanics.
-Similarly, in quantum mechanics, given an orthogonal subspace $A$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(A) = 2^{\dim(A)}$.
+In the continuous case, given a subset of phase space $A$ and $M_A$ the set of distributions with support within $A$, the maximum entropy is reached by the uniform distribution $\rho_A \in M_A$, which means we recover $\log_2 \capacity(M_A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
+Similarly, in quantum mechanics, given an orthogonal subspace $M$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(M) = 2^{\dim(M)}$.
 
@@
 In quantum mechanics, only orthogonal states are distinguishable, therefore the measure is non-additive over pure states.
-We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutual exclusiveness, is responsible for the failure of the additivity of the measure.
-Note that, since the fraction capacity depends only on the ensemble space, and not also on a chosen ensemble, the fraction capacity essentially characterizes the shape of the ensemble space itself.
+We see again how the lack of single decomposition of ensembles over pure states, which equates to the lack of mutual exclusivity, is responsible for the failure of the additivity of the measure.
+Note that, since the state capacity depends only on the ensemble space, and not also on a chosen ensemble, the state capacity essentially characterizes the shape of the ensemble space itself.
 
@@
 \textbf{Requirements for non-additive probability theory.} In a previous section we reviewed what types of objects we have in probability theory for classical ensembles: a sample space $X$, a probability measure $p$, random variables $F : X \to \mathbb{R}$, expectations as integrals of the probability measure $E[F] = \int_X F(x) dp$, a state counting measure $\mu$, the probability density as the Radon-Nikodym derivative $\frac{dp}{d\mu}$ and the entropy as $\int_X \rho \log \rho d\mu$.
-Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of the probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the states counting measure $\mu$, is there a generalization of all other elements?
+Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of the probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the state counting measure $\mu$, is there a generalization of all other elements?
 That is, can we find an integral and derivative operation over non-additive measures that generalizes to all physical theories?
@@
 The first difficulty, as we already pointed out, is that the ensemble space $\Ens$ may not have extreme points.
-What is the correct procedure to ``close'' the ensemble space and recover the space of extreme points is, for us, an open question.
+What the correct procedure to ``close'' the ensemble space and recover the space of extreme points is, for us, an open question.
 For now, we will assume $E$ is compact, it has a set of extreme points $X$ and the capacities have a unique restriction to the extreme points.
@@
 
-Recall that a statistical variable is a linear function $F : E \to \mathbb{R}$ defined over the ensemble space.
-Given that $F$ is a linear, its restriction on $X$ will determine the values for all other points of $E$.
+Recall that a statistical variable is an affine function $F : E \to \mathbb{R}$ defined over the ensemble space.
+Given that $F$ is affine, its restriction on $X$ will determine the values for all other points of $E$.
 Since the expectation is recovered through the integration of the variable over the probability density, we would like a non-additive integral $(E)\int_X$ such that
@@
 However, it does not work in the quantum case.
-We can show this by computing the Choquet integral for a two state quantum system and see that it does not recover the expectation of the statistical variables in the corresponding ensemble space.
+We can show this by computing the Choquet integral for a two-state quantum system and see that it does not recover the expectation of the statistical variables in the corresponding ensemble space.
 
@@
 On the left, the Bloch ball with a plane at constant $s$.
-On the right, the top case is when the value is greater than the maximum (empty set); the middle case is when the value is between the target point $\ens[o]$ and the maximum; the third case is when the value is below the target point.}
+On the right, the top case is when the value is greater than the maximum (empty set); the middle case is when the value is between the target point $\ens[o]$ and the maximum; the third case is when the value is below the target point.}\label{fig_ChoquetCuts}
 \end{figure}
@@
 A statistical variable $F$ (i.e.
-an observable), is a linear function of the ensemble space.
+an observable), is an affine function of the ensemble space.
 This means that two opposite points on the surface $\ens[a]$ and $\ens[b]$ (i.e.
@@
 The set $A(s) = \{x | F (x) \geq s\}$ corresponds to all the points of $X$, the surface of the sphere, that have a value of $F$ greater than a given $s$.
-Since $F$ is a linear function, this corresponds to the set of points above a given height.
+Since $F$ is an affine function, this corresponds to the set of points above a given height.
 So, all we need to do to calculate the integral is calculate the fraction capacity for sets of sphere $A(s)$ where the bottom is cut off at all possible values.
+See Fig.
+\ref{fig_ChoquetCuts}.
 
@@
 	\end{tikzpicture}
-	\caption{Geometric representation of the element $\ens[e]$ in $A(s)$ with the largest fraction of $\ens[o]$.}
+	\caption{Geometric representation of the element $\ens[e]$ in $A(s)$ with the largest fraction of $\ens[o]$.}\label{fig_ChoquetTop}
 \end{figure}
@@
 Since, for combinations along the same direction, $p$ increases as $\ens$ gets closer to $\ens[o]$ and as $\ens[c]$ gets further, $\ens[c]$ will be an extreme point on the lower hemisphere while $\ens$ will be on the disc at constant $s$, the lower side of the cap.
-Also note that the distance between $\ens[o]$ is the same for all extreme points, therefore, to maximize $p$, we simply need to minimize the distance between $\ens[o]$ and $\ens$.
-Therefore $\ens$ must be the intersection between the vertical axis and the disk, and $\ens[c]$ will coincide with $\ens[a]$.
+Also note that the distance to $\ens[o]$ is the same for all extreme points, therefore, to maximize $p$, we simply need to minimize the distance between $\ens[o]$ and $\ens$.
+Therefore $\ens$ must be the intersection between the vertical axis and the disk, and $\ens[c]$ will coincide with $\ens[a]$ as shown in Fig.
+\ref{fig_ChoquetTop}.
 The coefficient $p$ will be given by the ratio of the distance $\overline{\ens[o]\ens[a]}$ and $\overline{\ens\ens[a]}$.
-Since $F$ is a linearly functional, we have
+Since $F$ is an affine functional, we have
 \begin{equation}
@@
 
-To recap, we have
+To recap, as shown in Fig. \ref{fig_ChoquetIntegrand} we have
 \begin{equation}
@@
 	\end{tikzpicture}
-	\caption{The integrand $\frcap_{\ens[o]}(A(s))$ as a function of $s$.}
+	\caption{The integrand $\frcap_{\ens[o]}(A(s))$ as a function of $s$.}\label{fig_ChoquetIntegrand}
 \end{figure}
@@
 In terms of statistical variables, two mutually exclusive events can be assigned values independently.
-This, again, recovers how observable work in quantum mechanics, without mentioning quantum mechanics: only orthogonal states, that is mutually exclusive events, can be associated to measurement outcomes and be given a classical probability distribution.
+This, again, recovers how observables work in quantum mechanics, without mentioning quantum mechanics: only orthogonal states, that is mutually exclusive events, can be associated to different measurement outcomes and be given a classical probability distribution.
 The non-additivity of the fraction capacity and the non-arbitrariness of the statistical variables over the extreme points, then, must be linked to the shape of the ensemble space.
 
-Additionally, the non-additivity of the state capacity also depends on the degree of mutual exclusiveness of the events.
+Additionally, the non-additivity of the state capacity also depends on the degree of mutual exclusivity of the events.
 It is additive for mutually exclusive events and sub-additive otherwise.
@@
 One may be able to use the state capacity to characterize the degree of additivity of the fraction capacity.
-Since in non-additive measure theory, family of capacities are often characterized the type of non-additivity,\footnote{For example, a Sugeno $\lambda$-measure is a non-additive measure $\mu$ such that $\mu(A \cup B) = \mu(A) + \mu(B) + \lambda \mu(A)\mu(B)$ for some $\lambda$ over disjoint sets $A$ and $B$.
+Since in non-additive measure theory, families of capacities are often characterized by the type of non-additivity,\footnote{For example, a Sugeno $\lambda$-measure is a non-additive measure $\mu$ such that $\mu(A \cup B) = \mu(A) + \mu(B) + \lambda \mu(A)\mu(B)$ for some $\lambda$ over disjoint sets $A$ and $B$.
 If $\lambda = 0$, the measure is additive.} this may give a way forward.

### 2025-08-19 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: adding conclusion

@@
 
-We have seen that classical probability arises exactly when we are describing a set of ensembles that can always be decomposed in a unique set of mutually distinguishable cases.
-In quantum mechanics, instead, pure states
+We have seen how we can give a unified picture for how different measure theoretic tools can be used to represent states in physical theories.
+Standard probability theory can be understood as representing ensembles as mixtures of pure states, as extreme points.
+In particular, classical probability applies under the following equivalent conditions:
+\begin{enumerate}
+	\item the space of ensembles is a Choquet simplex
+	\item every probability distribution over the pure states identifies a unique ensemble
+	\item every ensembles is uniquely decomposable into pure states
+	\item sets of pure states identify mutually exclusive events
+	\item the fraction capacity of every ensemble is additive.
+\end{enumerate}
+Quantum mechanics fails to be characterized by classical probability precisely because it does not satisfies any of these conditions.
+We also saw that it is the embedding of any space of ensembles into a vector space that allows the use of quasi-probability, which is mandated by very minimal physical assumptions.
+However, quasi-probability representations are not unique, and therefore are not suitable as foundational descriptions.
+
+We have seen, instead, that non-additive measures can be created to generalize probability and state counting measure to the general case.
+Intuitively, the fraction capacity simply tells us how much of a particular ensemble can be understood as coming from a mixture of a set of other ensembles.
+The state capacity, instead, is highest exponential entropy achievable by mixing a given sets of ensembles, which reduces to the Liouville measure of classical mechanics and the dimensionality for quantum subspaces.
+
+In principle, these non-additive measures should allow a generalization of probability calculus that reduces to classical probability in the classical case.
+We have seen, however, that the two standard non-additive integrals (i.e.
+the Sugeno and the Choquet integral) are not suitable.
+It will likely need a broader collaboration to build these tools.
 
@@
 This article was made possible through the support of grant \#62847 from the John Templeton Foundation.
+We thank Vicen\c{c} Torra, Zuzana Ontkovi\v{c}ov\'{a} and Michel Grabisch for useful pointers and discussion.
 

### 2025-08-21 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
Squashed commit of the following:
commit 1b73e09bd459518237590eb79ee9634b9f4b9e01
Author: thrien <39382079+thrien@users.noreply.github.com>
Date:   Fri Aug 15 16:24:07 2025 -0400

    NonAdditiveProbability: Align dice example with book

    Instead of e_123456 = e_1 + e_23456 we split into e_3 + e_12456 like we do in the book.

commit 80a246abe6b3fbc1c477ff6866b3a2cd8eee94a5
Author: thrien <39382079+thrien@users.noreply.github.com>
Date:   Fri Aug 8 14:34:20 2025 -0400

    NonAdditiveProbability: Add old dice example

    Since I accidentally changed the previous example instead of adding a

@@
 
-In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts, [CITE - GPT, Ariel, Reginatto/Hall...] which stems from the underlying aim to characterize physical theories from an operational viewpoint.
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts \cite{plavala2023,ariano2017,caticha2019,hall2016}, [CITE - GPT, Ariel, Reginatto/Hall...] which stems from the underlying aim to characterize physical theories from an operational viewpoint.
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches \cite{ferrie10,mandel1995} are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 
@@
 
-\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasiprobability \cite{adhikary2020}.
+\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasiprobability \cite{ferrie2010,adhikary2020}.
 Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
@@
 Since we can write $\ens_{123456} = \frac{1}{6} \ens_{3} + \frac{5}{6} \ens_{12456}$, where $\ens_{12456}$ represents a uniform distribution over the five other outcomes, $1/6$ of $\ens_{123456}$ can be constructed from $A_3$, but no more.
+See Fig.
+\ref{fig_dice} for a visual representation.
 Similarly, if $A_{34} = \{\ens_{3},\ens_{4}\}$, we can write $\ens_{123456} = \frac{1}{3} \left(\frac{1}{2} \ens_3 + \frac{1}{2} \ens_4 \right)  + \frac{2}{3} \ens_{1256}$, so $1/3$ of $\ens_{123456}$ can be constructed from $A_{34}$, but no more.
@@
 This gives the basic intuition for our definitions.
+
+\begin{figure}[H]
+	\centering
+	\def\xstep{0.75em}
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
+	\end{tikzpicture}
+	$= \frac{1}{6}$
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(3, 1)};
+	\end{tikzpicture}
+	$+ \frac{5}{6}$
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/4) (4, 1/5) (5, 1/5) (6, 1/5)};
+	\end{tikzpicture}
+
+	\vspace{0.3cm}
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/5) (2, 1/5) (3, 1/5) (4, 1/5) (5, 1/5) (6, 1/5)};
+	\end{tikzpicture}
+	$= \frac{1}{3}$
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(3, 1/2) (4, 1/2)};
+	\end{tikzpicture}
+	$+ \frac{2}{3}$
+	\begin{tikzpicture}[y=3ex, x=\xstep]
+		\draw (0, 0) foreach \x in {1,...,6} { -- (\x, 0) -- (\x, -0.2) (\x, 0) } -- ++(1, 0);
+		\draw plot[ybar, bar width={\xstep / 2}] coordinates{(1, 1/4) (2, 1/4) (5, 1/4) (6, 1/4)};
+	\end{tikzpicture}
+	\caption{The fraction capacity is defined based on how much of a particular distribution can be decomposed in terms of other distributions.
+Here a uniform distribution over six cases can be seen as a mixture of different distributions.}\label{fig_dice}
+\end{figure}
 

### 2025-08-21 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex

@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches \cite{ferrie10,mandel1995} are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 
@@
 
-\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasiprobability \cite{ferrie2010,adhikary2020}.
+\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasiprobability \cite{adhikary2020}.
 Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.

### 2025-08-22 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Add citation for Monchietti

@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches \cite{monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 

### 2025-08-22 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex

--- 2025-08-21 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches \cite{ferrie10,mandel1995} are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches \cite{ferrie10,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 

### 2025-08-26 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: CHSH and Glauber reference
Added references to the CHSH inequality paper, a review of quasiprobability distributions by Glauber.

@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches \cite{ferrie10,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969}.[CITE] Quasiprobabilistic approaches \cite{ferrie10,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 
@@
 The most used is the Wigner function,[cite] that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through standard integration.
-While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.[cite] The key insight is that it is not negative probability that characterizes quasiprobability,\footnote{Some representations guarantee the quasiprobability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
+While the Wigner function represents the most known instance, there are different choices depending on the problem at hand\cite{cahill1969}.
+[cite] The key insight is that it is not negative probability that characterizes quasiprobability,\footnote{Some representations guarantee the quasiprobability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
 The quasiprobability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.

### 2025-08-26 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Cite Grabish, Pap, Torra, Sugeno
Just a note to remember citing Grabish, Pap, Torra, and Sugeno on non-additive measure theory.

@@
 Lastly, we lay down the requirements for a non-additive generalization of probability theory, which should be in principle possible.
-Since there is a wide literature on the subject, [cite] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
+Since there is a wide literature on the subject, [cite Grabish, Pap, TorraSugeno] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
 This means that more work will be needed to develop the appropriate tools.

### 2025-08-26 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Fix bibtex formatting

@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969}.[CITE] Quasiprobabilistic approaches \cite{ferrie10,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969}.[CITE] Quasiprobabilistic approaches \cite{ferrie2010,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 

### 2025-08-26 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Cite Peliti, Phelps, Mandel

@@
 
-We will first review how measure theory is used in classical probability [cite] and classical statistical mechanics[cite].
-We will connect this to Choquet theory[cite phelps], which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.
+We will first review how measure theory is used in classical probability [cite] and classical statistical mechanics \cite{peliti2024}.
+We will connect this to Choquet theory \cite{phelps2001}, which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.
 each ensemble can be represented by a probability measure over pure states).
@@
 We will then review some results from the use of quasiprobability measures.
-The most used is the Wigner function,[cite] that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through standard integration.
+The most used is the Wigner function \cite{mandel1995} that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through standard integration.
 While the Wigner function represents the most known instance, there are different choices depending on the problem at hand\cite{cahill1969}.

### 2025-08-26 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Use existing Peliti citation
As it turns out, there already exists a bibtex entry for Peliti's stat mech book. So I'm using that one instead. It seems more accurate anyway.
NonAdditiveProbability: Add proper citations
Add Grabisch to the bibliography.

@@
 
-We will first review how measure theory is used in classical probability [cite] and classical statistical mechanics \cite{peliti2024}.
+We will first review how measure theory is used in classical probability [cite] and classical statistical mechanics \cite{Peliti2011}.
 We will connect this to Choquet theory \cite{phelps2001}, which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.
@@
 While in classical statistical mechanics we start with a measure that quantifies the count of states and define the entropy using its logarithm, for the ensemble space we proceed in the opposite way: we start from the entropy and we define the count of states as the exponential of the entropy (i.e.
-the entropic volume).[cite Hall] From this, we can define a state capacity, which is again a sub-additive measure.
+the entropic volume)\cite{hall1999}.
+From this, we can define a state capacity, which is again a sub-additive measure.
 It reduces to the Liouville measure in the classical case and the dimensionality of subspaces in the quantum case.
@@
 Lastly, we lay down the requirements for a non-additive generalization of probability theory, which should be in principle possible.
-Since there is a wide literature on the subject, [cite Grabish, Pap, TorraSugeno] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
+Since there is a wide literature on the subject, \cite{pap2010,pap2013,nonadditive2014,grabisch2016}[cite Grabish, Pap, TorraSugeno] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
 This means that more work will be needed to develop the appropriate tools.
@@
 However, such a complete work would be effectively overwhelming both to produce and to read.
-The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source work.[cite our book]
+The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source work.\cite{aop-book}
 
@@
 Specifically, we are developing a general theory of ensemble spaces, the full treatment of which is beyond the scope of this work.
-Here we limit ourselves to giving only the initial premises and the final mathematical characterization, with more details in the appendix and in our ongoing work.[cite book]
+Here we limit ourselves to giving only the initial premises and the final mathematical characterization, with more details in the appendix and in our ongoing work.\cite{aop-book}
 
@@
 First, it needs to guarantee experimental verifiability.
-Mathematically, the ensemble space must be a $\mathsf{T}_0$ second countable topological space, where each open set represents an experimentally verifiable statement[cite book and kevin kelly].\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Second, it needs to allow statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.
+Mathematically, the ensemble space must be a $\mathsf{T}_0$ second countable topological space, where each open set represents an experimentally verifiable statement\cite{aop-book,kelly1996}.\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Second, it needs to allow statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.
 Mathematically, this constrains the ensemble space to be a convex set.\footnote{Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i=1}^{\infty} p_i \ens_i$ converges in the space or not.} Third, an entropy function must be defined that characterizes the variability of the elements within an ensemble.
@@
 The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
-fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\footnote{ \href{ https://link.springer.com/book/10.1007/978-3-319-03155-2}{cite} }
+fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\cite{nonadditive2014}
 

### 2025-08-27 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewing conclusion

--- 2025-08-26 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969}.[CITE] Quasiprobabilistic approaches \cite{ferrie2010,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics.[CITE] Quasiprobabilistic approaches [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 
@@
 The most used is the Wigner function,[cite] that represents each state as a signed probability measure over phase space, and expectations of observables are recovered through standard integration.
-While the Wigner function represents the most known instance, there are different choices depending on the problem at hand\cite{cahill1969}.
-[cite] The key insight is that it is not negative probability that characterizes quasiprobability,\footnote{Some representations guarantee the quasiprobability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
+While the Wigner function represents the most known instance, there are different choices depending on the problem at hand.[cite] The key insight is that it is not negative probability that characterizes quasiprobability,\footnote{Some representations guarantee the quasiprobability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
 The quasiprobability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
@@
 Lastly, we lay down the requirements for a non-additive generalization of probability theory, which should be in principle possible.
-Since there is a wide literature on the subject, [cite Grabish, Pap, TorraSugeno] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
+Since there is a wide literature on the subject, [cite] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
 This means that more work will be needed to develop the appropriate tools.
@@
 	\item every probability distribution over the pure states identifies a unique ensemble
-	\item every ensembles is uniquely decomposable into pure states
+	\item every ensemble is uniquely decomposable into pure states
 	\item sets of pure states identify mutually exclusive events
@@
 \end{enumerate}
-Quantum mechanics fails to be characterized by classical probability precisely because it does not satisfies any of these conditions.
+Quantum mechanics fails to be characterized by classical probability precisely because it does not satisfy any of these conditions.
 We also saw that it is the embedding of any space of ensembles into a vector space that allows the use of quasi-probability, which is mandated by very minimal physical assumptions.
@@
 
-We have seen, instead, that non-additive measures can be created to generalize probability and state counting measure to the general case.
+We have seen, instead, that non-additive measures can be created to generalize probability and state counting measure.
 Intuitively, the fraction capacity simply tells us how much of a particular ensemble can be understood as coming from a mixture of a set of other ensembles.
-The state capacity, instead, is highest exponential entropy achievable by mixing a given sets of ensembles, which reduces to the Liouville measure of classical mechanics and the dimensionality for quantum subspaces.
+The log of the state capacity, instead, is the highest entropy achievable by mixing a given set of ensembles, which reduces to the Liouville measure of classical mechanics and the dimensionality for quantum subspaces.
 

### 2025-08-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Added Wilce, Cahill, Mandel, Denneberg, and Grabisch

--- 2025-08-26 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts \cite{plavala2023,ariano2017,caticha2019,hall2016}, [CITE - GPT, Ariel, Reginatto/Hall...] which stems from the underlying aim to characterize physical theories from an operational viewpoint.
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts \cite{plavala2023,ariano2017,caticha2019,hall2016,wilce2025}, [CITE - GPT, Ariel, Reginatto/Hall...] which stems from the underlying aim to characterize physical theories from an operational viewpoint.
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
@@
 In some problems, especially in quantum optics, normal ordering and anti-normal ordering are more useful.
-The Glauber P [cite] representation and the Husimi Q [cite] representation work like the Wigner function except that the expectations of polynomials recover the respective order.
+The Glauber P \cite{cahill1969} representation and the Husimi Q \cite{mandel1995,cahill1969} representation work like the Wigner function except that the expectations of polynomials recover the respective order.
 In other circumstances, quasiprobability distributions over observables different from position and momentum are more useful.
@@
 
-\textbf{Ruling out Sugeno and Choquet integrals.} Given that extensive literature already exists for non-additive measure theory (cite), as it is used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence, we originally thought it would be just a matter of clarifying our definitions and connecting them to the existing literature.
+\textbf{Ruling out Sugeno and Choquet integrals.} Given that extensive literature already exists for non-additive measure theory \cite{denneberg1994non,grabisch2016}, as it is used in disparate contexts, such as game theory, economic decision theory, optimization and artificial intelligence, we originally thought it would be just a matter of clarifying our definitions and connecting them to the existing literature.
 However, things turned out to be more complicated.

### 2025-08-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Cite Fine

@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969}.[CITE] Quasiprobabilistic approaches \cite{ferrie2010,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969,fine1982}.[CITE] Quasiprobabilistic approaches \cite{ferrie2010,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 

### 2025-08-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Replace Ferrie with Zachos, add Kolmogorov

--- 2025-08-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969}.[CITE] Quasiprobabilistic approaches \cite{ferrie2010,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969,fine1982}.[CITE] Quasiprobabilistic approaches \cite{zachos2005,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 
@@
 
-We will first review how measure theory is used in classical probability [cite] and classical statistical mechanics \cite{Peliti2011}.
+We will first review how measure theory is used in classical probability \cite{kolmogorov2018} and classical statistical mechanics \cite{Peliti2011}.
 We will connect this to Choquet theory \cite{phelps2001}, which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.

### 2025-08-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: cleaning up citations

@@
 
-In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts \cite{plavala2023,ariano2017,caticha2019,hall2016,wilce2025}, [CITE - GPT, Ariel, Reginatto/Hall...] which stems from the underlying aim to characterize physical theories from an operational viewpoint.
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts \cite{plavala2023,ariano2017,caticha2019,hall2016,wilce2025}, which stems from the underlying aim to characterize physical theories from an operational viewpoint.
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
-A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969,fine1982}.[CITE] Quasiprobabilistic approaches \cite{zachos2005,mandel1995,monchietti2025} [CITE] are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
+A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969,fine1982}.
+Quasiprobabilistic approaches \cite{zachos2005,mandel1995,monchietti2025} are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
 
@@
 While the Wigner function represents the most known instance, there are different choices depending on the problem at hand\cite{cahill1969}.
-[cite] The key insight is that it is not negative probability that characterizes quasiprobability,\footnote{Some representations guarantee the quasiprobability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
+The key insight is that it is not negative probability that characterizes quasiprobability,\footnote{Some representations guarantee the quasiprobability to be non-negative.} but the lack of mutual exclusivity of the points of the space.
 The quasiprobability measures, in fact, are not defined over the extreme points, but over a suitable set $K$ such that the space of signed probability measures is isomorphic as a vector space to the one that embeds the ensemble space.
@@
 Lastly, we lay down the requirements for a non-additive generalization of probability theory, which should be in principle possible.
-Since there is a wide literature on the subject, \cite{pap2010,pap2013,nonadditive2014,grabisch2016}[cite Grabish, Pap, TorraSugeno] we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
+Since there is a wide literature on the subject, \cite{pap2010,pap2013,nonadditive2014,grabisch2016} we only show that the most used non-additive integrals, the Sugeno and Choquet integral, do not work for our purposes.
 This means that more work will be needed to develop the appropriate tools.
@@
 
-\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasiprobability \cite{ferrie2010,adhikary2020}.
+\textbf{Not about negative probability.} The inclusion of negative probability is often advertised as the main difference between classical probability and quasiprobability.
 Sometimes these negative probabilities are justified as arising from constraints on probability distributions defined on different contexts.
-This is a mischaracterization.
-For example, the Husimi Q representation provides us with a non-negative quasiprobability function.
+This is a mischaracterization.\cite{ferrie2010,adhikary2020} For example, the Husimi Q representation provides us with a non-negative quasiprobability function.
 That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.\footnote{Since the Husimi Q is defined as $Q_{\rho} ( \alpha ) = \frac{1}{\pi} \langle \alpha | \rho | \alpha \rangle$, it is proportional to the probability of measuring a coherent (i.e.

### 2025-08-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: further cleanup

@@
 
-%TODO: literature search on use of signed measure in QM. ( e.g. \href{https://arxiv.org/pdf/2302.00118}{this paper})
-
 \section{Introduction}
 
-In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts \cite{plavala2023,ariano2017,caticha2019,hall2016,wilce2025}, which stems from the underlying aim to characterize physical theories from an operational viewpoint.
+In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts \cite{plavala2023,wilce2025,ariano2017,caticha2019,hall2016}, which stems from the underlying aim to characterize physical theories from an operational viewpoint.
 One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
@@
 \section{Quasiprobability}
-
-%Goal of this section: quasiprobability representations are allowed by the embedding in a topological vector space.
-Choice of representations are guided by what observables are interesting for a specific problem.
-
-%Examples: Wigner, Husimi, Glauber P, SIC-POVM (space not made of spectra of operators), Kirkwood-Dirac (two-time correlation function - observables not necessarily taken at the same time )
-
-%Things to say: some people think about it as weakining axioms (cite papers on negative probability).
-It's more than that.
-Points describe events that are not necessarily mutually exclusive.
-TODO: example of two Wigner function/Husimi Q would be better that have no overlap but they are not orthogonal and/or two functions that have overlap but are orthogonal.
-
-%Thing to say: all these representations are useful because they preserve the linear structure, meaning that convex combinations of ensembles are convex combinations of quasiprobability; expectation values are linear.
-Check, but it should be true.
-
-% 1 describe quasiprobability approaches; go through examples; point out that the quasiprobability can be negative; that one can recover expectations;
 
@@
 Therefore, at a conceptual level, probabilities can only give us a patchwork representation of ensembles.
-Some authors embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.
-[CITATION NEEDED] However, there is a different route.
+Some authors embrace this view, by characterizing states as the collection of all probability distributions of all possible results, each distribution providing partial information.\cite{plavala2023,wilce2025} However, there is a different route.
 

### 2025-09-03 Articles/2025-HilbertSpaceResponse/main.tex
HilberSpaceResponse: as on arxiv

\documentclass[pdflatex,sn-mathphys-num]{sn-jnl}
\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

\begin{document}

\title[Article Title]{Response to ``Are Hilbert Spaces Unphysical? Hardly, My Dear!''}

\author*[1]{\fnm{Gabriele} \sur{Carcassi}}\email{carcassi@umich.edu}

\author{\fnm{Robert} \sur{Rozite}}

\author[1]{\fnm{Christine A.} \sur{Aidala}}

\affil[1]{\orgdiv{Physics Department}, \orgname{University of Michigan}, \orgaddress{\street{450 Church St}, \city{Ann Arbor}, \state{MI} \postcode{48109}, \country{United States}}}

%\affil*[2]{\orgdiv{Physics Department}, \orgname{University of Wisconsin-Madison}, \orgaddress{\street{1150 University Ave}, \city{Madison}, \postcode{53706}, \state{WI}, \country{United States}}}

%orcid for Gabriele Carcassi: https://orcid.org/0000-0002-1071-6251

%orcid for Robert Rozite: https://orcid.org/0009-0007-6133-998X

\abstract{A recent criticism of our paper ``The unphysicality of Hilbert spaces'' by Nivaldo Lemos refutes our central argument that a state with finite expectation value can be mapped to a state with infinite expectation value by a coordinate transformation. By conflating coordinate transformation with change of basis in quantum mechanics, Lemos argues that expectation values are invariant under change of variables. In the present work, we clarify the distinction between coordinate transformation and change of basis, and rebut Lemos' main argument.}

\keywords{Quantum foundations, Hilbert spaces}
\maketitle

%%do we need a section title here?

In our original paper \cite{carcassi}, we claim that Hilbert spaces are problematic because they allow coordinate transformations that map states with finite expectation values to ones with infinite expectations. In his recent criticism \cite{lemos}, Nivaldo Lemos sums up our approach: ``through a change of variables, which is a unitary transformation representing a change of basis, the state $\psi$ in which the expectation value of any positive even power $X^n$ of the position operator is finite has been turned into a state $\phi$ in which these expectation values are infinite.'' He continues to show that expectation values are invariant under changes of basis, and that, therefore, our proposed coordinate transformations cannot change expectation values. The conflation of coordinate transformations and change of basis lies at the core of Lemos' central argument, and will be shown to be inconsistent with established literature. More specifically, Lemos says we reached an ``unsound conclusion'' because we ``interpreted the change of variables as a unitary transformation that changes the state vectors but leaves the observables alone''. This is exactly our procedure, and is consistent with an active transformation as described in various textbooks \cite{gottfried2, ballentine, binney}. From Shankar's ``Principles of Quantum Mechanics'' \cite{shankar}:

\begin{quote}
    Suppose we subject all the vectors \(|V\rangle\) in a space to a unitary transformation
\begin{align}
    |V \rangle \rightarrow U |V \rangle
\end{align}
Under this transformation, the matrix elements of any operator \(\Omega\) are modified as follows: 
\begin{align}
    \langle V' | \Omega | V \rangle \rightarrow \langle UV' | \Omega | UV \rangle = \langle V' | U^\dag \Omega U |V\rangle
\end{align}
It is clear that the same change would be effected if we left the vectors alone and subjected all operators to the change
\begin{align}
    \Omega \rightarrow U^\dag \Omega U
\end{align}
The first case is called an active transformation and the second a passive transformation. The present nomenclature is in reference to the vectors: they are affected in an active transformation and left alone in the passive case. The situation is exactly the opposite from the point of view of the operators.

Later we will see that the physics in quantum theory lies in the matrix elements of operators, and that active and passive transformations provide us with two equivalent ways of describing the same physical transformation.

\end{quote}

Coordinate transformations and changes of variables are similar, if not equivalent, concepts that can be interpreted as physical transformations. A simple example is spatial translation, which is carried out as a change of variables: $x \rightarrow x + a$. This example is well established in the literature \cite{gottfried1, binney}, and, intuitively, does not preserve the expectation value of position.

Still, Lemos insists that a coordinate transformation is a unitary transformation representing a change of basis. Given a unitary transformation, \(U\), a change of basis is carried out by transforming states as \(\phi' = U \phi\) and operators as \(O' = UOU^\dag\). Unitarity requires \(UU^\dag = U^\dag U = I\). As such, expectation values are invariant under a change of basis: \(\langle \phi' | O' | \phi' \rangle = \langle \phi | O | \phi \rangle\). Lemos purports that we have failed to transform the position operator in addition to the state, which would be true if we were performing a change of basis. Considering our coordinate transformation as an active transformation, \textit{not} a change of basis, his argument quickly falls apart. Furthermore, he claims ``every time-independent unitary transformation in the Hilbert space of states is a change of basis''. While unitary transformations are used to perform changes of basis, they are more generally used to transform the state of a physical system to another state, without necessarily preserving expectation values. An insistence that states and operators both be transformed by a unitary transformation would preclude us from studying changes of reference frame and time evolution which, through relativity, are intimately linked.

Despite over a hundred years of quantum physics, it is disheartening that a criticism conflating coordinate transformation with change of basis can pass through peer review uncontested. It is our hope that the present work not only rebuts Lemos' central argument, but also clarifies the procedure and intuition behind physical transformations and change of basis in quantum mechanics.

%Lemos further claims that ``any change of basis or change of representation can always be regarded as an active unitary transformation in Hilbert space that changes vectors and operators.'' We find this language to be inconsistent with the literature.

\section*{Acknowledgments}
This paper is part of the ongoing Assumptions of Physics project~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived. One of the authors of the original paper did not have time to contribute to this response. The other authors thank Robert Rozite for taking the lead in preparing this response.

\section*{Conflict of Interest}
The corresponding author states that there is no conflict of interest.
\section*{Data Availability Statement}
Not applicable.
\bibliography{bibliography}

\end{document}

### 2025-09-05 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: Dumped axioms and propositions
I dumped all the axioms with their justifications and key propositions for state and fraction capacity from the book into the appendix.
I excluded the mathSection environment, but kept the axiom and justification environment.
This is just the raw material that needs to be trimmed down.

There are some inconsistencies between the book and the paper:
\size vs. \fraction
\chull vs. \hull

@@
 \newcommand\fraction{\mathrm{frac}}
+\newcommand\size{\mathrm{frac}}
 \newcommand\frcap{\mathrm{fcap}}
+\newcommand\fcap{\mathrm{fcap}}
+\newcommand\chull{\mathrm{chull}}
 
@@
 
+\appendix
+\section*{Axioms}
+\begin{axiom}[Axiom of ensemble] 
+	The state of a system is given by an \textbf{ensemble}, which represents all possible preparations of equivalent systems prepared according to the same procedure.
+The set of all possible ensembles for a particular system is an \textbf{ensemble space}.
+Formally, an ensemble space is a $\mathsf{T}_0$ second countable topological space where each element is called an ensemble.
+\end{axiom}
+
+\begin{justification}
+	In experimental settings, preparation procedures never prepare a system exactly in the same configuration.
+Experimental results, then, are always in terms of statistical preparations and statistical measurements.
+A physical theory, then, must be able to talk about the possible statistical descriptions possible by the theory.
+States, then, can be understood as ensembles, idealized statistical descriptions, as those are what is connected to experimental practice.
+
+	Equivalently, reproducibility is a basic requirement of a physical theory.
+A physical law, then, must be understood as describing a relationship that always exists whenever the same set of circumstances is replicated.
+Given that we need to always be able to replicate those circumstances ``one more time'', the relationship is about countably infinite preparations and results: ensembles.
+Therefore, to the extent that physics is about reproducible experimental results, the basic theoretical description of a system is in terms of ensembles.
+This justifies the use of ensembles as the fundamental object to describe the state of a system.\footnote{Note that reproducibility also already implies that all properties that characterize an ensemble must be relative to the procedure.
+If the properties depended, for example, on absolute space or absolute time, then different practitioners would not be able to prepare the same ensemble.}
+
+	Ensembles are experimentally defined objects, and therefore they are possibilities of an experimental domain.
+Therefore an ensemble space is a $\mathsf{T}_0$ second countable topological space where each element is an ensemble and the topology is induced by the verifiable statements.
+\end{justification}
+
+\begin{defn}
+	Given a real number $p \in [0,1]$, its complement is defined as $\bar{p} = 1-p$.
+\end{defn}
+
+\begin{axiom}[Axiom of mixture]\label{pm_es_axiomMixture}
+	The statistical mixture of two ensembles is an ensemble.
+Formally, an ensemble space $\Ens$ is equipped with an operation $+ : [0, 1] \times \Ens \times \Ens \to \Ens$ called \textbf{mixing}, noted with the infix notation $p \ens[a] + \bar{p} \ens[b]$, with the following properties:
+	\begin{itemize}
+		\item \textbf{Continuity}: the map $+(p, \ens[a], \ens[b])  \to p \ens[a] + \bar{p} \ens[b]$ is continuous (with respect to the product topology of $[0, 1] \times \Ens \times \Ens$)
+		\item \textbf{Identity}: $1 \ens[a] + 0 \ens[b] = \ens[a]$
+		\item \textbf{Idempotence}:  $p \ens[a] + \bar{p} \ens[a] = \ens[a]$ for all $p \in [0,1]$
+		\item \textbf{Commutativity}: $p \ens[a] + \bar{p} \ens[b] = \bar{p} \ens[b] + p \ens[a]$ for all $p \in [0,1]$
+		\item \textbf{Associativity}: $p_1 \ens_1 + \bar{p}_1\left(\frac{1-p_1-p_3}{\bar{p}_1}\ens_2 + \frac{p_3}{\bar{p}_1}\ens_3\right) =  \bar{p}_3\left(\frac{p_1}{\bar{p}_3} \ens_1 +  \frac{1-p_1-p_3}{\bar{p}_3}\ens_2\right) + p_3 \ens_3$ where $p_1, p_3 \in [0,1)$ and $p_1 + p_3 \leq 1$
+	\end{itemize}
+\end{axiom}
+
+\begin{justification}
+	This axiom captures the ability to create a mixture merely by selecting between the output of different processes.
+Let $\ens_1$ and $\ens_2$ be two ensembles that represent the output of two different processes $P_1$ and $P_2$.
+Let a selector $S_p$ be a process that outputs two symbols, the first with probability $p$ and the second with probability $\bar{p}$.
+Then we can create another process $P$ that, depending on the selector, outputs either the output of $P_1$ or $P_2$.
+All possible preparations of such a procedure will form an ensemble.
+Therefore we are justified in equipping an ensemble space with a mixing operation that takes a real number from zero to one, and two ensembles.
+
+	Given that mixing represents an experimental relationship, and all experimental relationships must be continuous in the natural topology, mixing must be a continuous function.
+Note that $p$ is a continuously ordered quantity, where no value is perfectly experimentally verifiable, and therefore the natural topology is the one of the reals.
+This justifies continuity.
+
+	If $p=1$, the output of $P$ will always be the output of $P_1$.
+This justifies the identity property.
+If $P_1$ and $P_2$ are the same process, then the output of $P$ will always be the output of $P_1$.
+This justifies the idempotence property.
+The order in which the processes are given does not matter as long as the same probability is matched to the same process.
+The process $P$ is identical under permutation of $P_1$ and $P_2$.
+This justifies commutativity.
+If we are mixing three processes $P_1$, $P_2$ and $P_3$, as long as the final probabilities are the same, it does not matter if we mix $P_1$ and $P_2$ first or $P_2$ and $P_3$.
+This justifies associativity.
+\end{justification}
+
+\begin{coro}
+	An ensemble space is a convex space.
+\end{coro}
+
+\begin{proof}
+	The properties of the axiom of mixture match the basic definition of convex spaces, which can be taken from \href{https://ncatlab.org/nlab/show/convex+space}{ncatlab} and \href{https://arxiv.org/abs/0903.5522}{this excellent paper}.
+The notation and terminology will be slightly different to better map to physics ideas.
+\end{proof}
+
+\begin{axiom}[Axiom of entropy]
+	Every element of the ensemble is associated with an \textbf{entropy} which quantifies the variability of the preparations of the ensemble.
+Formally, an ensemble space $\Ens$ is equipped with a function $S : \Ens \to \mathbb{R}$, defined up to a positive multiplicative constant representing the unit numerical value.
+The entropy has the following properties:
+	\begin{itemize}
+		\item \textbf{Continuity}\footnote{Currently, we are imposing that the entropy is continuous.
+There may be a chance that this requirement is redundant, as strict concavity and the upper variability bound may already impose this.
+We have found proofs that show that real valued convex/concave functions of real values are continuous.
+These proofs fail at the extreme points, but the upper variability bound may fix this.
+Another open question is whether differentiability is also an independent requirement.}
+		\item \textbf{Strict concavity}: $S(p\ens[a] + \bar{p} \ens[b]) \geq p S(\ens[a]) + \bar{p} S(\ens[b])$ with the equality holding if and only if $\ens[a] = \ens[b]$
+		\item \textbf{Upper variability bound}: there exists a universal function $I(p_1, p_2)$ (i.e.
+the same for all ensemble spaces) such that $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p, \bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])$; if the equality holds, $\ens[a]$ and $\ens[b]$ are \textbf{mutually exclusive} or \textbf{orthogonal}, noted $\ens[a] \ortho \ens[b]$
+		\item \textbf{Mixtures preserve orthogonality}:\footnote{It is unclear whether ``mixtures preserve orthogonality'' is an independent axiom.
+Intuitively, the following argument tells us that it is.
+Take the 2 dimensional simplex (i.e.
+a triangle) that represents a classical discrete probability space over three elements.
+Take the standard entropy, which will satisfy ``mixtures preserve orthogonality''.
+This is because the middle point has entropy $\log 3$.
+We can imagine redefining the entropy so that it is a little bit lower in the center but it is unchanged on the sides.
+However, one needs to provide an actual example and show that it satisfies all axioms.
+It may also be that only one direction is an independent axiom.
+That is, that orthogonality with the components implies orthogonality with the mixtures.
+This is what does not hold for separateness.} $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$
+	\end{itemize}
+\end{axiom}
+
+\begin{justification}
+	The entropy quantifies the variability of the instances within an ensemble.
+Since the ensemble represents a collection of preparations of equivalent systems, and since each instance will in general be potentially different, it is legitimate to ask how much variability there is among the different instances.
+We are assuming that the entropy is a quantity (i.e.
+a linearly ordered property), meaning that it is always meaningful to tell whether one ensemble has more variability than another.
+If this is the case, the later requirements of continuity and strict concavity will force the entropy to be a real valued quantity.
+This is because the variability will change under statistical mixtures, and since statistical mixtures are performed with real valued coefficients, the variability will have to be a real valued quantity.
+Whether it is conceptually possible to have a characterization of variability that is not linearly ordered but still have a meaningful connection with statistical mixing is an open question, therefore we are not able to fully justify entropy's linear ordering at this time.
+Therefore the linear ordering of the entropy should be considered an assumption.
+Provided that assumption, we are justified to assume the existence of a real valued function that returns the entropy, a measure of variability of the ensemble.
+
+	Since the entropy is a real valued quantity, it will have a corresponding unit.
+This unit is independent from all other units, and therefore the overall structure of the ensemble space must be independent of this choice.
+Mathematically, the physical dimension of the unit is not captured, just its numeric value.
+A change of unit may change the numeric value by a multiplicative constant.
+Since variability is an ordered quantity, we want the change of units to respect the ordering and therefore it should be a positive multiplicative constant.
+This justifies that the entropy function is defined up to a positive multiplicative constant.
+
+	Note that the additivity of the entropy over independent systems fixes the absolute scale.
+If $S_{AB} = S_A + S_B$, in fact, one can't rescale all three terms by an additive factor and preserve the relationship.
+
+	The variability, in the end, will have physical consequences, and it will therefore be measurable, thus experimentally verifiable: it will have to be a topologically continuous function.
+Moreover, small changes in the ensemble should produce small changes in the variability, which justifies analytical continuity.
+We are therefore justified to assume continuity of the entropy.
+
+	Suppose we have two ensembles and we perform a statistical mixture.
+There are going to be three sources of variability: the two ensembles and the random choice at every instance.
+The total contribution from the original ensembles will be the average variability of the original ensembles.
+This is increased by the variability introduced by the random choice, which is always a positive contribution.
+Therefore the final variability cannot be less than the average of the original ensembles.
+That is, $S(p\ens[a] + \bar{p} \ens[b]) \geq p S(\ens[a]) + \bar{p} S(\ens[b])$.
+If we are mixing an ensemble with itself, this is equivalent to just choosing from the original ensemble, therefore the variability will not increase.
+Conversely, if the variability stays the same, it means that the random choice does not increase the variability, and therefore we must be choosing between equivalent ensembles.
+Therefore we are justified to assume that entropy is strictly concave.
+
+	On the other hand, the variability cannot increase arbitrarily during mixture.
+The maximum variability will be given when the two ensembles are mutually exclusive, when an instance of the first ensemble cannot be produced by the second ensemble.
+That is, a single instance is enough to determine whether we have the first ensemble or the second.
+In this case, the variability is increased by the variability of the random choice, which must depend only on the mixture coefficient, and not the nature of the ensembles themselves.
+That is, $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p, \bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])$ is the upper variability bound, which is saturated if and only if the $\ens[a]$ and $\ens[b]$ are mutually exclusive.
+The actual function $I$ is left unspecified and, as we show in proposition \ref{pm_es_entropyUnique}, it will correspond to the Shannon entropy as it is the only indicator of variability that will satisfy the axiom of entropy.
+This justifies the upper variability bound.
+
+	Now suppose ensemble $\ens[a]$ is mutually exclusive from both $\ens[b]$ and $\ens[c]$.
+That is, an instance of $\ens[a]$ cannot ever be produced by either $\ens[b]$ or $\ens[c]$.
+Then an instance of $\ens[a]$ cannot be produced by a mixture of $\ens[b]$ and $\ens[c]$, since ultimately a mixture of $\ens[b]$ and $\ens[c]$ will return an instance of one of the two.
+Therefore $\ens[a]$ is mutually exclusive from any mixture of $\ens[b]$ and $\ens[c]$.
+The argument works in reverse as well: if an instance of $\ens[a]$ cannot be produced by a mixture of $\ens[b]$ and $\ens[c]$, then it cannot be produced by either.
+This justifies mixtures preserve orthogonality.
+\end{justification}
+
+\section{Propositions}
+\subsection{State capacity}
+\begin{prop}
+	The state capacity is a set function that is
+	\begin{enumerate}
+		\item non-negative: $\capacity(A) \in [0, +\infty]$
+		\item monotone: $A \subseteq B \implies \capacity(A) \leq \capacity(B)$
+		\item subadditive: $\capacity(A \cup B) \leq \capacity(A) + \capacity(B)$
+		\item additive over orthogonal sets: $A \ortho B \implies \capacity(A \cup B) = \capacity(A) + \capacity(B)$ 
+	\end{enumerate}
+\end{prop}
+
+\begin{proof}
+	1.
+The state capacity takes a subset of $\Ens$ and returns a real value and is therefore a set function.
+The exponential can only return non-negative values, therefore the state capacity of a set is non-negative.
+
+	2. The $\hull$ is a monotone function. The image of a set through a map is a monotone function. The supremum is a monotone function. Therefore the state capacity is a monotone function.
+
+	3.
+Let $A, B \subseteq \Ens$ and let $\ens = p \ens[a] + \bar{p} \ens[b]$ for some $p \in [0,1]$, $\ens[a] \in A$ and $\ens[b] \in B$.
+By \ref{pm_es_exponentialEntropySubadditivity} and the definition of state capacity, $2^{S(\ens)} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])} \leq \capacity(A) + \capacity(B)$.
+Since this is true for any finite mixture of elements of $A$ and $B$, it will be true for any element of $\hull(A \cup B)$.
+Consequently, the supremum of the exponential entropy cannot exceed the sum of the state capacities.
+Therefore $\capacity(A \cup B) \leq \capacity(A) + \capacity(B)$ (i.e.
+the state capacity is subadditive).
+
+	4.
+Let $A, B \subseteq \Ens$ be two orthogonal subsets.
+Let $\{\ens[a]_i\} \subseteq A$ be a sequence of ensembles such that $2^{S(\ens[a]_i)} \to \capacity(A)$ and let $\{\ens[b]_i\} \subseteq B$ be a sequence of ensembles such that $2^{S(\ens[b]_i)} \to \capacity(B)$.
+Consider $\ens_i = p_i \ens[a]_i + \bar{p_i} \ens[b]_i$ where $p_i = \frac{2^{S(\ens[a]_i)}}{2^{S(\ens[a]_i)} + 2^{S(\ens[b]_i)}}$.
+Then, by \ref{pm_es_exponentialEntropySubadditivity}, $2^{S(\ens_i)} = 2^{S(\ens[a]_i)} + 2^{S(\ens[b]_i)}$.
+This means that $2^{S(\ens_i)} \to \capacity(A) + \capacity(B)$.
+Therefore $\capacity(A \cup B)$ must be at least $\capacity(A) + \capacity(B)$.
+Combining with the previous result, $\capacity(A \cup B) = \capacity(A) + \capacity(B)$.
+Therefore the state capacity, as a set function, is additive over orthogonal sets of ensembles.
+\end{proof}
+
+\begin{prop}
+	The state capacity is continuous from above and below.
+\end{prop}
+
+\begin{proof}
+	Let $\{A_i\} \subseteq \Sigma_{\Ens}$ be an increasing sequence.
+Given that the state capacity is monotone, $\{\capacity(A_i)\}$ is a an increasing sequence and it will converge in the extended real line.
+Let $A = \lim\limits_{i \to \infty} A_i$.
+We have $A = \cup A_i$ and $A_i \subseteq A$ for all $i$.
+Since state capacity is monotone, $\capacity(A_i) \leq \capacity(A)$ for all $i$ and therefore $\lim\limits_{i \to \infty} \capacity(A_i) \leq \capacity(A)$.
+Given any $\ens[a] \in \chull(A)$, we can find a sequence $\ens[a]_i \in \chull(A_i)$ such that $\ens[a]_i \to \ens[a]$.
+Because of the continuity of the entropy, $S(\ens[a]_i) \to S(\ens[a])$.
+This means we cannot have $k$ such that $\capacity(A) > k > \capacity(A_i)$ for all $i$, and therefore $\lim\limits_{i \to \infty} \capacity(A_i) = \capacity(A)$.
+
+	The same argument works for decreasing sequences by inverting the order.
+\end{proof}
+
+\subsection{Fraction capacity}
+\begin{coro}
+	The fraction capacity uniquely identifies an ensemble. That is, let $\ens[a], \ens[b] \in \Ens$ such that $\ens[a] \neq \ens[b]$. Then $\fcap_{\ens[a]} \neq \fcap_{\ens[b]}$.
+\end{coro}
+
+\begin{proof}
+	Note that $\ens = 1 \ens[a] + 0 \ens[b]$ if and only if $\ens = \ens[a]$.
+Therefore, let $\ens[a], \ens[b] \in \Ens$ such that $\ens[a] \neq \ens[b]$.
+We have $\fcap_{\ens[a]}(\{\ens[a]\}) = 1$ and $\fcap_{\ens[b]}(\{\ens[a]\}) \neq 1$.
+Which means $\fcap_{\ens[a]} \neq \fcap_{\ens[b]}$.
+\end{proof}
+
+\begin{prop}
+	The fraction capacity for an ensemble is a set function that is
+	\begin{enumerate}
+		\item non-negative and unit bounded: $\fcap_{\ens}(A) \in [0,1]$
+		\item monotone: $A \subseteq B \implies \fcap_{\ens}(A) \leq \fcap_{\ens}(B)$
+		\item subadditive: $\fcap_{\ens}(A \cup B) \leq \fcap_{\ens}(A) + \fcap_{\ens}(B)$
+		\item continuous from below: $\fcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \fcap_{\ens}(A_i)$ for any increasing sequence $\{A_i\}$
+		\item continuous from above: $\fcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \fcap_{\ens}(A_i)$ for any decreasing sequence $\{A_i\}$
+	\end{enumerate}
+\end{prop}
+
+\begin{proof}
+	1.
+The fraction capacity takes a subset of $\Ens$ and returns a real value and is therefore a set function.
+The mixture coefficients are real values between zero and one.
+The supremum of a set of numbers between zero and one is between zero and one, and therefore the fraction capacity for an ensemble is non-negative and unit bounded.
+
+	2. The $\hull$ is a monotone function. The image of a set through a map is a monotone function. The supremum is a monotone function. Therefore the fraction capacity is a monotone function.
+
+	3.
+Let $A, B \subseteq \Ens$ and let $p \in [0,1]$ such that $\ens = p \ens_1 + \bar{p} \ens_2$ for some $\ens_1 \in \hull(A \cup B)$ and $\ens_2 \in \Ens$.
+Since $\ens_1 \in \hull(A \cup B)$, we can write $\ens_1 = \lambda \ens[a] + \bar{\lambda} \ens[b]$ for some $\lambda \in [0,1]$, $\ens[a] \in A$ and $\ens[b] \in B$.
+Therefore we have $\ens = p \lambda \ens[a] + p \bar{\lambda} \ens[b] + \bar{p} \ens_2$.
+By the definition of fraction capacity, we must have $p\lambda \leq \fcap_{\ens}(A)$ and $p\bar{\lambda} \leq \fcap_{\ens}(B)$, therefore $p = p\lambda + p\bar{\lambda} \leq \fcap_{\ens}(A) + \fcap_{\ens}(B)$.
+	Since $\fcap_{\ens}(A \cup B)$ is the supremum for a set of $p$s for which the expression always holds, we have $\fcap_{\ens}(A \cup B) \leq \fcap_{\ens}(A) + \fcap_{\ens}(B)$.
+The fraction capacity is subadditive.
+
+	4.
+Let $\{A_i\} \subseteq \Sigma_{\Ens}$ be an increasing sequence.
+That is $A_i \subseteq A_{i+1}$ for all $i$.
+Given that the fraction capacity is monotone, $\{\fcap_{\ens}(A_i)\}$ is an increasing sequence.
+Note that $\fcap_{\ens}(A_i) \leq 1$ for all $i$, therefore the sequence will have an upper bound which will coincide with the limit.
+Let $A = \lim\limits_{i \to \infty} A_i$.
+We have $A = \bigcup A_i$ and $A_i \subseteq A$ for all $i$.
+Since fraction capacity is monotone, we have $\fcap_{\ens}(A_i) \leq \fcap_{\ens}(A)$ for all $i$ and therefore $\lim\limits_{i \to \infty} \fcap_{\ens}(A_i) \leq \fcap_{\ens}(A)$.
+Suppose $\lim\limits_{i \to \infty} \fcap_{\ens}(A_i) < \fcap_{\ens}(A)$.
+Then there would be an $\ens[a] \in \hull(A)$ such that $\size_{\ens}(\ens[a]) > \size_{\ens}(\ens[b])$ for all $\ens[b] \in \hull(A_i)$ for all $i$.
+But if $\ens[a] \in \hull(A)$, then it can be written as a finite combination of finitely many elements of $A$.
+Then we will be able to find a $k$ such that $A_k$ will contain all the elements, and therefore $\ens[a] \in A_k$.
+But this means $\ens[a] \leq \fcap_{\ens}(A_k)$ which is a contradiction.
+Therefore $\lim\limits_{i \to \infty} \fcap_{\ens}(A_i) = \fcap_{\ens}(A)$.
+
+	5. The previous argument works in the same way for decreasing sequences by inverting the order.
+\end{proof}
+
 \end{document}

### 2025-09-06 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
OrderTheory: reviewing a bit

@@
 \begin{abstract}
-This note presents a condensed summary of order theory which can function as a crash course, refresher and/or reference.
-Bare minima are meant to give a rough overview, by no means complete, of the subject to the intellectually curious, particularly in the context of foundational questions in physics.
+A summary overview of order theory, which can be used as a reference, refresher or to be aware of what one does not know.
 \end{abstract}
@@
 
-The mathematical structures studied by order theory appear in different contexts, making the topic relevant to those interested in foundational questions.
-Core structures in fields such mathematical logic, topology and measure theory, can be characterized and understood in purely order theoretic terms; characterizing number types (i.e.
-integers, rational and reals) as different order types makes a better link to physics; physical concepts (e.g.
+Order theory studies partially ordered sets (POSETs): sets equipped with a notion of order.
+These structures are pervasive in mathematics as play a role in logic, computer science, topology, measure theory, group theory, and beyond.
+Even number types types (i.e.
+integers, rational and reals) can be characterized purely on their order structure.
+A as different order types makes a better link to physics; physical concepts (e.g.
 part-whole inclusion) are also naturally captured with order theory.
@@
 For more details on order theory, see for example \cite{davey2002introduction, pinter2014book}.
-
-\emph{This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.}
 
@@
 
-Mathematically, the following is the most convenient and common way to define order relationships.
-
 \begin{defn}[Partial orders and poset]
@@
 	\end{description}
-	A \textbf{partially ordered set}, or \textbf{poset}, is a tuple $(X, \leq)$ consisting on a set and a partial order.
+	A \textbf{partially ordered set}, or \textbf{poset}, is a tuple $(X, \leq)$ consisting of a set and a partial order.
 \end{defn}
@@
 \begin{remark}
-	To give a sense of the broad range of applicability: sets are ordered by set inclusion ($\{0, 1\} \subseteq \{0, 1, 2\}$); statements are ordered by implication/narrowness (\statement{Snoopy is a dog} $\narrower$ \statement{Snoopy is a mammal}); systems are ordered by is-a-part-of (`piston' $\subseteq$ `engine' $\subseteq$ `car').
+	To give a sense of the broad range of applicability: sets are ordered by set inclusion ($\{0, 1\} \subseteq \{0, 1, 2\}$); statements are ordered by implication/narrowness/specificity (\statement{Snoopy is a dog} $\narrower$ \statement{Snoopy is a mammal}); systems are ordered by is-a-part-of (`piston' $\subseteq$ `engine' $\subseteq$ `car').
 \end{remark}
@@
 Now define $\lesssim$ on $X_{/\sim}$ such that $a_{/\sim} \lesssim b_{/\sim}$ if $a \lesssim b$.
-This is a preorder.
+This is a partial order.
 \end{prop}
@@
 Simple orders \marginleft{Hasse diagram}  can be depicted using \textbf{Hasse diagrams}.
-Each element in the poset is represented by a vertex.
+Each element of the poset is represented by a vertex.
 A line connects upwards an element to its immediate successor, so that lowest elements are at the bottom.
@@
 \begin{defn}
-	A \textbf{monotonic function} \marginleft{Monotonic function} (also called \textbf{monotone}, \textbf{increasing} or \textbf{order preserving}) is a map between two posets $(X, \leq_X)$ and $(Y, \leq_Y)$ that respects the order.
+	An \textbf{increasing function} \marginleft{Increasing function} (also called \textbf{order preserving}) is a map between two posets $(X, \leq_X)$ and $(Y, \leq_Y)$ that preserves the order.
 That is, a map $f : X \to Y$ such that $a \leq_X b \implies f(a) \leq_Y f(b)$.
+It is \textbf{strictly increasing} if $a \leq_X b \implies f(a) <_Y f(b)$.
 \end{defn}
@@
 \begin{defn}
-	An \textbf{order embedding} \marginleft{Order embedding} is a monotonic function that respects the order in both directions. That is, a map $f : X \to Y$ such that $a \leq_X b \iff f(a) \leq_Y f(b)$.
+	A \textbf{decreasing function} \marginleft{Decreasing function} (also called \textbf{order reversing}) is a map between two posets $(X, \leq_X)$ and $(Y, \leq_Y)$ that reverses the order.
+That is, a map $f : X \to Y$ such that $a \leq_X b \implies f(b) \leq_Y f(a)$.
+It is \textbf{strictly decreasing} if $a \leq_X b \implies f(b) <_Y f(a)$.
+\end{defn}
+
+\begin{defn}
+	A \textbf{monotonic function} \marginleft{Monotonic function} (also called \textbf{monotone}) is a map that is either increasing or decreasing.
+\end{defn}
+
+\begin{defn}
+	An \textbf{order embedding} \marginleft{Order embedding} is a monotonic function where the image have the same ordering. That is, a map $f : X \to Y$ such that $a \leq_X b \iff f(a) \leq_Y f(b)$.
 \end{defn}
@@
 bijective monotonic function whose inverse is monotonic).
-Two posets are \textbf{order isomorophic} if there exist an order isomoprhism between them.
+Two posets are \textbf{order isomorphic} (or have the same \textbf{order type}) if there exists an order isomorphism between them.
 \end{defn}
+
+\subsection{Filters and ideals AFTER}
 
@@
 
+\begin{remark}
+	The upper/lowers bound of $A$ are not necessarily in $A$.
+\end{remark}
+
 \begin{defn}[Supremum and infimum]
@@
 \begin{defn}[Join and meet]
-	The \textbf{join} \marginleft{Join and meet: $\vee, \wedge$} of two elements $a, b \in X$ or of a set $A$, noted $a \vee b$ and $\bigvee A$, is their supremum if it exists.
-The \textbf{meet} of two elements $a, b \in X$, noted $a \wedge b$, is their infimum, if it exists.
+	The \textbf{join} \marginleft{Join and meet: $\vee, \wedge$} of two elements $a, b \in X$ or of a set $A \subseteq X$, noted $a \vee b$ and $\bigvee A$, is their supremum if it exists.
+The \textbf{meet} of two elements $a, b \in X$ or of a set $A \subseteq X$, noted $a \wedge b$ and $\bigwedge A$, is their infimum, if it exists.
 \end{defn}
@@
 		A complete lattice is \textbf{infinitely distributive} if the same holds over infinite operations.
-		\item[Complemented.] A bounded lattice is \textbf{complemented} if every element has a unique complement.
+		\item[Complemented.] A bounded lattice is \textbf{complemented} if every element has a complement. It is uniquely complement if the complement is unique.
+		\item[Ortho-complemented.] A bounded lattice is \textbf{complemented} if every element has a unique complement.
 	\end{description}

### 2025-09-11 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
OrderTheory: more work on the bare minimum

@@
 \end{defn}
+
+\begin{defn}[Maximal subsets]
+	A linearly \marginleft{Maximal subset} ordered subset (i.e. a chain) is \textbf{maximal} if it is not a scrict subset of another linearly ordered subset.
+\end{defn}
+
+\begin{prop}[Kuratowski's Lemma or Hausdorff maximal principle]
+	Every linearly ordered subset is the subset of some maximal linearly ordered subset.
+\end{prop}
+
+\begin{prop}
+	Kuratowski's Lemma is equivalent to the axiom of choice.
+\end{prop}
 
@@
 \begin{defn}[Upper and lower bounds]
-	Let $P$ be an \marginleft{Upper/Lower bounds} ordered set and let $A \subseteq P$.
-An \textbf{upper bound} (or \textbf{lower bound}) is an element $p \in P$ such that $a \leq p$ (or $p \leq a$) for all $a \in A$.
+	Let $X$ be an \marginleft{Upper/Lower bounds} ordered set and let $A \subseteq X$.
+An \textbf{upper bound} (or \textbf{lower bound}) is an element $x \in X$ such that $a \leq x$ (or $x \leq a$) for all $a \in A$.
 \end{defn}
 
-\begin{remark}
-	The upper/lowers bound of $A$ are not necessarily in $A$.
-\end{remark}
+\begin{prop}[Zorn's lemma]
+	Every non-empty partially ordered set for which every linearly ordered subset has an upper bound contains at least one maximal element.
+\end{prop}
 
@@
 		\item[Completeness.] A lattice is \textbf{complete} if all subsets have joins and meet, not just finite ones.
+		\item[Modularity.] A lattice is \textbf{modular} if $a \leq b$ implies $a \vee (c \wedge b) = (a \vee c) \wedge b$
 		\item[Distributivity.] A lattice is \textbf{distributive} if it satisfies
@@
 		\item[Complemented.] A bounded lattice is \textbf{complemented} if every element has a complement. It is uniquely complement if the complement is unique.
-		\item[Ortho-complemented.] A bounded lattice is \textbf{complemented} if every element has a unique complement.
+		\item[Orthocomplemented.] A bounded lattice is \textbf{orthocomplemented} if for every element $a$ there is a complement $a^{\bot}$ such that $a^{\bot\bot} = a$ and if $a \leq b$ then $b^{\bot} \leq a^{\bot}$
+		\item[Orthomodular.] A lattice is \textbf{orthomodular} if it is orthocomplemented and if $a \leq b$ implies $a \vee (a^{\bot} \wedge b) = (a \vee a^{\bot}) \wedge b$ (i.e.
+modularity holds at least of orthocomplements)
 	\end{description}
 \end{defn}
+
+\begin{prop}
+	Every distributive lattice is a modular. Every complete lattice is bounded.
+\end{prop}
 
@@
 \end{remark}
-
-\begin{prop}[Zorn's lemma]
-	Every non-empty partially ordered set for which every linearly ordered subset has an upper bound contains at least one maximal element.
-\end{prop}
-
-\begin{defn}[Maximal subsets]
-	A linearly \marginleft{Maximal subset} ordered subset is \textbf{maximal} if it is not a scrict subset of another linearly ordered subset.
-\end{defn}
-
-\begin{prop}[Kuratowski's Lemma]
-	Every linearly ordered subset is the subset of some maximal linearly ordered subset.
-\end{prop}
 

### 2025-09-17 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
OrderTheory: added closures and intersection structures

@@
 
+\section{Closures and intersection structures}
+
+\begin{defn}
+	Given a poset $(X, \leq)$ a \textbf{closure operator} is a map $c : X \to X$ that satisfies the following for all $a,b \in X$:
+	\begin{enumerate}
+		\item \textbf{extensive}: $a \leq c(a)$
+		\item \textbf{increasing}: $a \subseteq b \implies c(a) \subseteq c(b)$
+		\item \textbf{idempotent}: $c(c(a)) = c(a)$.
+	\end{enumerate}
+\end{defn}
+
+\begin{defn}
+	Given a set $S$, an \textbf{intersection structure}, or $\bigcap$-structure, on $S$ is a family $X \subseteq 2^{S}$ of subsets of $S$ that is closed under arbitrary intersection.
+That is, $\bigcap A_i \in X$ for every non-empty family $A_i \in X$.
+A topped intersection structure on $S$ is an intersection structure such that $S \in X$.
+\end{defn}
+
+\begin{prop}
+	Let $C : 2^{S} \to 2^{S}$ be a closure operator.
+Then the set of closed sets $X = \{ A \subset S \, | \, C(A) = A\}$ is a topped intersection structure.
+Conversely, Let $X$ be a topped intersection structure, $C(A) = \bigcap \{B \in X \, | \, | A \subset B \}$ is a closure operator.
+\end{prop}
+
 \bibliographystyle{plain}

### 2025-09-21 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
OrderTheory: major improvements

--- 2025-09-11 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
@@
 }
+\usepackage[all,pdf]{xy}
 \frenchspacing
@@
 
-Order theory studies partially ordered sets (POSETs): sets equipped with a notion of order.
-These structures are pervasive in mathematics as play a role in logic, computer science, topology, measure theory, group theory, and beyond.
-Even number types types (i.e.
-integers, rational and reals) can be characterized purely on their order structure.
-A as different order types makes a better link to physics; physical concepts (e.g.
+Order theory studies partially ordered sets (posets): sets equipped with a notion of order.
+These structures are pervasive in mathematics as they play a role in logic, computer science, topology, measure theory, group theory, and beyond.
+Even number types (i.e.
+integers, rational and reals) can be characterized purely by their order structure.
+Some physical concepts (e.g.
 part-whole inclusion) are also naturally captured with order theory.
@@
 For more details on order theory, see for example \cite{davey2002introduction, pinter2014book}.
-
-\section{TBD}
-
-* fixpoint theorems
-* Zorn's lemma
-
-* Galois connection
-
-? graded posets
 
@@
 
-\begin{remark}
-	To give a sense of the broad range of applicability: sets are ordered by set inclusion ($\{0, 1\} \subseteq \{0, 1, 2\}$); statements are ordered by implication/narrowness/specificity (\statement{Snoopy is a dog} $\narrower$ \statement{Snoopy is a mammal}); systems are ordered by is-a-part-of (`piston' $\subseteq$ `engine' $\subseteq$ `car').
-\end{remark}
+Example of posets: family of sets ordered by set inclusion ($\{x, y\} \subseteq \{x, y, z\}$); statements ordered by implication/narrowness/specificity (\statement{Snoopy is a dog} $\narrower$ \statement{Snoopy is a mammal}); systems ordered by is-a-part-of (`piston' $\subseteq$ `engine' $\subseteq$ `car'); numbers ordered by magnitude ($2 \leq 3$).
+
+\begin{defn}[Subposet]
+	Given a poset $(X, \leq)$, a \textbf{subposet} is a poset $(Y, \leq_Y)$ where $Y \subset X$ and $\leq_Y = \leq |_Y$.
+\end{defn}
+
+\marginright{
+$$
+\def\arl{\ar@{-}}
+\xymatrix@R=10pt@C=4pt{
+	& \{x,y,z\}\arl[dl]\arl[d]\arl[dr] & \\
+	\{x,y\}\arl[d]\arl[dr] & \{x,z\}\arl[dl]|\hole\arl[dr]|\hole & \{y,z\}\arl[dl]\arl[d] \\
+	\{x\}\arl[dr]   & \{y\}\arl[d]   & \{z\}\arl[dl] \\
+	& \{\varnothing\} \\
+}
+$$
+}
+Simple orders \marginleft{Hasse diagram}  can be depicted using \textbf{Hasse diagrams}.
+Each element of the poset is represented by a vertex.
+A line connects upwards an element to its immediate successor, so that lowest elements are at the bottom.
 
@@
 
-The more intuitive notion of `before' and `after' would not include the equal. The two notions can be shows to be equivalent.
+The more intuitive notion of `before' and `after' would not include the equal. The two notions can be shown to be equivalent.
 
@@
 
-A more loose idea of order is a preorder, for which two different elements can be in the same order position. For example, the relationship has-no-fewer-elements for set is a preorder.
+\begin{defn}[Predecessors and successors]
+	A \textbf{predecessor} \marginleft{Predecessor, successor} (or \textbf{successor}) of an element $a \in X$ is an element that comes strictly before (or strictly after) $a$.
+That is, it is an element $b \in X$ such that $b < a$ (or $a < b$).
+A predecessor (or successor) is  \textbf{immediate} if there are no elements in between.
+That is, $b < a$ (or $a < b$) and there is no $c \in X$ such that $b < c < a$  (or $a < c < b$).
+\end{defn}
+
+\begin{defn}[Dense order]
+	An order is \textbf{dense} \marginleft{Dense order} if for any two elements $a < b$ we can always find a third element $c \in X$ between the two $a < c < b$.
+\end{defn}
+
+\begin{defn}[Discrete order]
+	A \textbf{discrete order} \marginleft{Discrete order} is an order in which every element that has a predecessor/successor has an immediate predecessor/successor.
+\end{defn}
+
+Note that discrete order may also mean an order in which every element is only related to itself (i.e. the partial order is an equivalence relation).
+
+A more loose idea of order is a preorder, for which two different elements can be in the same order position.
 
@@
 
+Examples of preorders: multi-dimensional objects compared by a single variable; sets compared by their cardinality; sets of integers compared by the sum of the elements.
+
 \begin{prop}
@@
 
-\begin{defn}[Predecessors and successors]
-	A \textbf{predecessor} \marginleft{Predecessor, successor} (or \textbf{successor}) of an element $a \in X$ is an element that comes strictly before (or strictly after) $a$.
-That is, it is an element $b \in X$ such that $b < a$ (or $a < b$).
-A predecessor (or successor) is  \textbf{immediate} if there are no elements in between.
-That is, $b < a$ (or $a < b$) and there is no $c \in X$ such that $b < c < a$  (or $a < c < b$).
-\end{defn}
-
-\marginright{TODO: Hesse diagram}
-
-Simple orders \marginleft{Hasse diagram}  can be depicted using \textbf{Hasse diagrams}.
-Each element of the poset is represented by a vertex.
-A line connects upwards an element to its immediate successor, so that lowest elements are at the bottom.
-
-\begin{defn}[Dense order]
-	An order is \textbf{dense} \marginleft{Dense order} if for any two elements $a < b$ we can always find a third element $c \in X$ between the two $a < c < b$.
-\end{defn}
-
 \subsection{Morphisms}
@@
 That is, a map $f : X \to Y$ such that $a \leq_X b \implies f(a) \leq_Y f(b)$.
-It is \textbf{strictly increasing} if $a \leq_X b \implies f(a) <_Y f(b)$.
+It is \textbf{strictly increasing} if $a <_X b \implies f(a) <_Y f(b)$.
 \end{defn}
@@
 That is, a map $f : X \to Y$ such that $a \leq_X b \implies f(b) \leq_Y f(a)$.
-It is \textbf{strictly decreasing} if $a \leq_X b \implies f(b) <_Y f(a)$.
+It is \textbf{strictly decreasing} if $a <_X b \implies f(b) <_Y f(a)$.
 \end{defn}
@@
 \begin{defn}
-	An \textbf{order embedding} \marginleft{Order embedding} is a monotonic function where the image have the same ordering. That is, a map $f : X \to Y$ such that $a \leq_X b \iff f(a) \leq_Y f(b)$.
+	An \textbf{order embedding} \marginleft{Order embedding} is a monotonic function where the image has the same ordering. That is, a map $f : X \to Y$ such that $a \leq_X b \iff f(a) \leq_Y f(b)$.
 \end{defn}
@@
 	An \textbf{order isomorphism} \marginleft{Order isomorphism} is an invertible embedding (i.e.
-bijective monotonic function whose inverse is monotonic).
+bijective increasing function whose inverse is increasing).
 Two posets are \textbf{order isomorphic} (or have the same \textbf{order type}) if there exists an order isomorphism between them.
-\end{defn}
-
-\subsection{Filters and ideals AFTER}
-
-\begin{defn}
-	Let $X$ be a poset.
-\marginleft{Up-sets and \\ down-sets} An \textbf{up-set}, or \textbf{increasing set}, or \textbf{order filter}, is a subset $A \subseteq X$ that contains all the successors of its elements.
-That is, if $a \in A$ then for all $b \in X$ such that $a \leq b$ we have $b \in A$.
-Dually, a \textbf{down-set}, or \textbf{decreasing set}, or \textbf{order ideal}, is a subset $A \subseteq X$ that contains all the predecessors of its elements.
-That is, if $a \in A$ then for all $b \in X$ such that $b \leq a$ we have $b \in A$.
 \end{defn}
@@
 
+\begin{prop}
+	Zorn's lemma is equivalent to the axiom of choice.
+\end{prop}
+
 \begin{defn}[Supremum and infimum]
@@
 \begin{defn}[Lattice - order theoretic]
-	A \textbf{lattice} \marginleft{Lattice} is a partially ordered sets in which every pair of elements has both a supremum/join and an infimum/meet.
+	A \textbf{lattice} \marginleft{Lattice} is a partially ordered set in which every pair of elements has both a supremum/join and an infimum/meet. A \textbf{sublattice} is a subposet of a lattice.
 \end{defn}
@@
 \begin{prop}[Connecting lemma]
-	If $X$ is a lattice then $a \leq b$, $a \vee b = b$ and $a \wedge b = a$ are equivalent.
+	If $X$ is a lattice then $a \leq b \iff a \vee b = b \iff a \wedge b = a$.
 \end{prop}
@@
 	\begin{description}
-		\item[Completeness.] A lattice is \textbf{complete} if all subsets have joins and meet, not just finite ones.
-		\item[Modularity.] A lattice is \textbf{modular} if $a \leq b$ implies $a \vee (c \wedge b) = (a \vee c) \wedge b$
-		\item[Distributivity.] A lattice is \textbf{distributive} if it satisfies
+		\item[Completeness] A lattice is \textbf{complete} if all subsets have joins and meet, not just finite ones.
+		\item[Modularity] A lattice is \textbf{modular} if $a \leq b$ implies $a \vee (c \wedge b) = (a \vee c) \wedge b$
+		\item[Distributivity] A lattice is \textbf{distributive} if it satisfies
 		\begin{align*}
@@
 		A complete lattice is \textbf{infinitely distributive} if the same holds over infinite operations.
-		\item[Complemented.] A bounded lattice is \textbf{complemented} if every element has a complement. It is uniquely complement if the complement is unique.
-		\item[Orthocomplemented.] A bounded lattice is \textbf{orthocomplemented} if for every element $a$ there is a complement $a^{\bot}$ such that $a^{\bot\bot} = a$ and if $a \leq b$ then $b^{\bot} \leq a^{\bot}$
-		\item[Orthomodular.] A lattice is \textbf{orthomodular} if it is orthocomplemented and if $a \leq b$ implies $a \vee (a^{\bot} \wedge b) = (a \vee a^{\bot}) \wedge b$ (i.e.
+		\item[Complemented] A bounded lattice is \textbf{complemented} if every element has a complement. It is \textbf{uniquely complemented} if the complement is unique.
+		\item[Orthocomplemented] A bounded lattice is \textbf{orthocomplemented} if for every element $a$ there is a complement $a^{\bot}$ such that $a^{\bot\bot} = a$ and if $a \leq b$ then $b^{\bot} \leq a^{\bot}$.
+		\item[Orthomodular] A lattice is \textbf{orthomodular} if it is orthocomplemented and if $a \leq b$ implies $a \vee (a^{\bot} \wedge b) = (a \vee a^{\bot}) \wedge b$ (i.e.
 modularity holds at least of orthocomplements)
@@
 \end{defn}
+\marginright{
+	$$
+	\def\arl{\ar@{-}}
+	\xymatrix@R=7pt@C=4pt{
+		& M_3 & \\
+		 & \top\arl[dl]\arl[d]\arl[dr] & \\
+		a\arl[dr] & b\arl[d] & c\arl[dl] \\
+		& \bot \\
+	}
+	\xymatrix@R=2pt@C=4pt{
+		& N_5 & \\
+		& \top\arl[dl]\arl[ddr] & \\
+		a\arl[dd] &  & \\
+		 &  & c\arl[ddl] \\
+		b\arl[dr] &  &  \\
+		& \bot \\
+}
+	$$
+}
+\begin{prop}
+	Every distributive lattice is a modular. Every complete lattice is bounded.
+\end{prop}
 
 \begin{prop}
-	Every distributive lattice is a modular. Every complete lattice is bounded.
+	A lattice is distributive if an only if it does not contain an $M_3$ (diamond) or $N_5$ (pentagon) sublattice.
 \end{prop}
@@
 \begin{remark}
-	The completion is constructed by finding pairs $(A, B)$ of subsets $A, B \subseteq X$ such that $A$ is the set of lower bounds for $B$ and $B$ is the set of lower bounds of $A$.
+	The completion is constructed by finding pairs $(A, B)$ of subsets $A, B \subseteq X$ such that $A$ is the set of lower bounds for $B$ and $B$ is the set of upper bounds of $A$.
 For the rationals $\mathbb{Q}$, these pairs correspond to Dedekind cuts, and the completion of $\mathbb{Q}$ as a lattice is the lattice of the reals $\mathbb{R}$.
@@
 
-\section{Algebras and orders}
+\subsection{Filters, ideals and other subsets}
+
+\begin{defn}
+	Let $X$ be a poset.
+\marginleft{Upward and \\ downward closed} A subset $A \subseteq X$ is \textbf{upward closed} (or \textbf{downward closed}), if it contains all elements greater (or less) than of its elements.
+That is, if $a \in A$, then $b \in A$ for all $b \in X$ such that $a \leq b$ (or $b \leq a$).
+\end{defn}
+
+\begin{defn}
+	Let $X$ be a poset.
+\marginleft{Upward and \\ downward \\  directed} A subset $A \subseteq X$ is \textbf{upward directed} (or \textbf{downward directed}), if any finite set has an upper (or lower) bound.
+That is, if $a,b \in A$, then we can find $c \in A$ such that $a \leq c$ and $b \leq c$ (or $c \leq a$ and $c \leq b$).
+\end{defn}
+
+\begin{defn}
+	Let $X$ be a poset.
+\marginleft{Filters and \\ ideals} A subset $A \subseteq X$ is a \textbf{filter} if it is upward closed and downward directed.
+It is a \textbf{proper filter} if $A \neq X$.
+It is a \textbf{principal filter}, noted $\uparrow a$ if it contains and only contains all elements greater than $a \in X$.
+Conversely, $A$ is an \textbf{ideal} if it is downward closed and upward directed.
+It is a \textbf{proper ideal} if $A \neq X$.
+It is a \textbf{principal ideal}, noted $\downarrow a$ if it contains and only contains all elements less than $a \in X$.
+\end{defn}
+
+\begin{defn}
+	A filter (or ideal) \marginleft{Maximal \\ filters \\ and ideals} is \textbf{maximal} if it is proper and no other filter (or ideal) contains it.
+A maximal filter is also called an \textbf{ultrafilter}.
+\end{defn}
+
+\section{Lattices as algebras}
+
+Lattices can be equivalently defined as algebraic structures.
 
@@
 	\begin{description}
-		\item[Associativity.] $(a \vee b) \vee c = a \vee (b \vee c)$
+		\item[Associativity] $(a \vee b) \vee c = a \vee (b \vee c)$
 		and $(a \wedge b) \wedge c = a \wedge (b \wedge c)$
-		\item[Commutativity.] $a \vee b = b \vee a$
+		\item[Commutativity] $a \vee b = b \vee a$
 		and $a \wedge b = b \wedge a$
-		\item[Idempotency.] $a \vee a = a$
+		\item[Idempotency] $a \vee a = a$
 		and $a \wedge a = a$
-		\item[Absorption.] $a \vee (a \wedge b) = a$
+		\item[Absorption] $a \vee (a \wedge b) = a$
 		and $a \wedge (a \vee b) = a$
@@
 \begin{prop}
-	Every order lattice is an algebraic lattice and every algebraic lattice is an order lattice where $a \leq b$ if $a \vee b = b$.
+	Every order lattice is an algebraic lattice and every algebraic lattice is an order lattice where $a \leq b$ if and only if $a \vee b = b$.
 \end{prop}
@@
 	\begin{description}
-		\item[Lattice.] $(X, \vee, \wedge)$ is a lattice
-		\item[Distributivity.] $a \vee (b \wedge c) = (a \vee b) \wedge (a \vee b)$
+		\item[Lattice] $(X, \vee, \wedge)$ is a lattice
+		\item[Distributivity] $a \vee (b \wedge c) = (a \vee b) \wedge (a \vee b)$
 and $a \wedge (b \vee c) = (a \wedge b) \vee (a \wedge b)$
-		\item[Identity.] $a \vee \bot = a$
+		\item[Identity] $a \vee \bot = a$
 		and $a \wedge \top = a$
-		\item[Complementation.] $a \vee \neg a = \top$
+		\item[Complementation] $a \vee \neg a = \top$
 		and $a \wedge \neg a = \bot$
@@
 	All Heyting algebras are distributive lattices.
-Every Boolean algebra is a Heyting algebra where the pseudo-complement coincide with the complement.
+Every Boolean algebra is a Heyting algebra where the pseudo-complement coincides with the complement.
 If a Heyting algebra is complemented then it is a Boolean algebra.
@@
 
-\section{Representation as sets}
+\section{Set representation of lattices}
 
@@
 		\item[Meet] represented by the intersection
-		\item[Pseudo-complement] represented by the set complement.
+		\item[Complement] represented by the set complement.
 	\end{description}
@@
 \begin{prop}[Lattice of subgroups]
-	Given a group, the set of its subgroups, as a collection of sets ordered by inclusion, is a lattice following correspondence:
+	Given a group, the set of its subgroups, as a collection of sets ordered by inclusion, is a lattice with the following correspondence:
 	\begin{description}
@@
 \begin{prop}[Lattice of subspaces]
-	Given a vector space, the set of its subspaces, as a collection of sets ordered by inclusion, is a complemented lattice following correspondence:
+	Given a vector space, the set of its subspaces, as a collection of sets ordered by inclusion, is a complemented lattice with the following correspondence:
 	\begin{description}
@@
 		\item[Meet] represented by the intersection
-		\item[Complement] represented by the orthogonal subspace.
+		\item[Complement] represented by the orthogonal subspace (for inner product spaces).
 	\end{description}
@@
 	Note how the order theoretic operations (i.e.
-join, meet, complement, ...), except for Boolean algebras, do not always map to the set theoretic operations (i.e.
-intersection, union, complement, ...).
-This creates confusion when one tries to attach intuition from one to the other.
+join, meet, complement), except for Boolean algebras, do not always map to the set theoretic operations (i.e.
+intersection, union, complement).
+This may be confusing when one tries to attach intuition from one to the other.
 
@@
 However, the join fails to map to the union.
-Logically, the disjunction of \statement{the spin of system was prepared in the z+ direction} and \statement{the spin of system was prepared in the z- direction} is \statement{the spin of system was prepared in either the z+ or z- direction} while the join in terms of subspaces is \statement{the spin of system was prepared in any direction}.
+Logically, the disjunction of \statement{the spin of the system was prepared in the z+ direction} and \statement{the spin of the system was prepared in the z- direction} is \statement{the spin of the system was prepared in either the z+ or z- direction} while the join in terms of subspaces is \statement{the spin of the system was prepared in any direction}.
 
@@
 
+All number types we use to represent the value of physical quantities can be characterized by their ordering relationship.
+This includes their topology, as it is the order topology in all cases.
+Note that the complex numbers do not fit this pattern.
+
 \begin{prop}
-	Every linearly ordered set $X$ in which each element has one predecessor and one successor is order isomorphic to the integers $\mathbb{Z}$.
+	Every discretely linearly ordered set with a lower bound and no upper bound is order isomorphic to the naturals $\mathbb{N}$.
 \end{prop}
@@
 \begin{prop}
+	Every unbounded discretely linearly ordered set is order isomorphic to the integers $\mathbb{Z}$.
+\end{prop}
+
+\begin{prop}[Cantor's isomorphism theorem]
 	 Every unbounded countable dense linearly ordered set is order isomorphic to the rationals $\mathbb{Q}$.
@@
 
+\begin{prop}[Cantor's characterization of the real line]
+	Every complete unbounded dense linearly ordered set which admits a countable dense subset is order isomorphic to the reals $\mathbb{R}$.
+\end{prop}
+
+\section{Closures and intersection structures}
+
+\begin{defn}
+	Given a poset $(X, \leq)$ a \textbf{closure operator} is a map $cl : X \to X$ that satisfies the following for all $a,b \in X$:
+	\begin{enumerate}
+		\item \textbf{extensive}: $a \leq cl(a)$
+		\item \textbf{increasing}: $a \leq b \implies cl(a) \leq cl(b)$
+		\item \textbf{idempotent}: $cl(cl(a)) = cl(a)$.
+	\end{enumerate}
+\end{defn}
+
+Examples of closure operators: the ceiling function that returns the smallest integer greater than a given real number; the span operator over a vector space; the topological closure.
+
+\begin{defn}
+	Given a set $S$, an \textbf{intersection structure}, or $\bigcap$-structure, on $S$ is a family $X \subseteq 2^{S}$ of subsets of $S$ that is closed under arbitrary intersection.
+That is, $\bigcap A_i \in X$ for every non-empty family $A_i \in X$.
+A topped intersection structure on $S$ is an intersection structure such that $S \in X$.
+\end{defn}
+
+Examples of topped intersection structures: the lattice of subspaces of a given vector space, the lattice of subgroups of a group, the family of closed subsets of a topological space.
+
 \begin{prop}
-	Every complete unbounded dense linearly ordered set which admits a countable dense subset is order isomorphic to the reals $\mathbb{R}$.
+	Let $C : 2^{S} \to 2^{S}$ be a closure operator.
+Then the set of closed sets $X = \{ A \subseteq S \, | \, C(A) = A\}$ is a topped intersection structure.
+Conversely, Let $X$ be a topped intersection structure, $C(A) = \bigcap \{B \in X \, | \, A \subseteq B \}$ is a closure operator.
 \end{prop}

### 2025-09-21 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
OrderTheory: fixed left margin notes

@@
 \begin{defn}[Lattice - order theoretic]
-	A \textbf{lattice} \marginleft{Lattice} is a partially ordered set in which every pair of elements has both a supremum/join and an infimum/meet. A \textbf{sublattice} is a subposet of a lattice.
+	A \textbf{lattice} \marginleft{Lattice \\ (order)} is a partially ordered set in which every pair of elements has both a supremum/join and an infimum/meet.
+A \textbf{sublattice} is a subposet of a lattice.
 \end{defn}
@@
 \begin{defn}[Lattice - algebraic]
-	An algebraic \marginleft{Lattice (algebraic)} structure $(X, \vee, \wedge)$ consisting of a set $X$ and two binary operations $\vee$ and $\wedge$ is an \textbf{(algebraic) lattice} if it satisfies the following properties:
+	An algebraic \marginleft{Lattice \\(algebra)} structure $(X, \vee, \wedge)$ consisting of a set $X$ and two binary operations $\vee$ and $\wedge$ is an \textbf{(algebraic) lattice} if it satisfies the following properties:
 	\begin{description}
@@
 \begin{defn}[Boolean algebra]
-	An algebraic \marginleft{Boolean algebra} structure $(X, \vee, \wedge, \neg, \bot, \top)$ is a \textbf{Boolean algebra} if it satisfies the following properties:
+	An algebraic \marginleft{Boolean \\ algebra} structure $(X, \vee, \wedge, \neg, \bot, \top)$ is a \textbf{Boolean algebra} if it satisfies the following properties:
 	\begin{description}
@@
 \begin{defn}
-	A \textbf{Heyting algebra} \marginleft{Heyting algebra} is a bounded lattice for which the relative pseudo-complement always exists.
+	A \textbf{Heyting algebra} \marginleft{Heyting \\ algebra} is a bounded lattice for which the relative pseudo-complement always exists.
 \end{defn}
@@
 \begin{defn}
-	Given a poset $(X, \leq)$ a \textbf{closure operator} is a map $cl : X \to X$ that satisfies the following for all $a,b \in X$:
+	Given a poset $(X, \leq)$ \marginleft{Closure} a \textbf{closure operator} is a map $cl : X \to X$ that satisfies the following for all $a,b \in X$:
 	\begin{enumerate}
@@
 \begin{defn}
-	Given a set $S$, an \textbf{intersection structure}, or $\bigcap$-structure, on $S$ is a family $X \subseteq 2^{S}$ of subsets of $S$ that is closed under arbitrary intersection.
+	Given a set $S$, \marginleft{Intersection \\structure} an \textbf{intersection structure}, or $\bigcap$-structure, on $S$ is a family $X \subseteq 2^{S}$ of subsets of $S$ that is closed under arbitrary intersection.
 That is, $\bigcap A_i \in X$ for every non-empty family $A_i \in X$.

### 2025-09-21 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
OrderTheory: final review

@@
 Even number types (i.e.
-integers, rational and reals) can be characterized purely by their order structure.
+integers, rationals and reals) can be characterized purely by their order structure.
 Some physical concepts (e.g.
@@
 \begin{defn}[Subposet]
-	Given a poset $(X, \leq)$, a \textbf{subposet} is a poset $(Y, \leq_Y)$ where $Y \subset X$ and $\leq_Y = \leq |_Y$.
+	Given a poset $(X, \leq)$, a \textbf{subposet} is a poset $(Y, \leq_Y)$ where $Y \subseteq X$ and $\leq_Y = \leq |_{Y \times Y}$.
 \end{defn}
@@
 \begin{defn}[Linear order]
-	A \textbf{linear order}, \marginleft{Linear order,\\ chain} or \textbf{total order}, is a partial order where every pair is ordered.
+	A \textbf{linear order}, \marginleft{Linear order,\\ chain, \\anti-chain} or \textbf{total order}, is a partial order where every pair of elements is comparable.
 That is, given $a, b \in X$ either $a \leq b$ or $b \leq a$.
 A \textbf{chain} typically refers to a subset of a partial order that is linearly ordered.
+An \textbf{anti-chain} is a subset of a partial order in which elements are only comparable to themselves.
 \end{defn}
@@
 \begin{defn}[Maximal subsets]
-	A linearly \marginleft{Maximal subset} ordered subset (i.e. a chain) is \textbf{maximal} if it is not a scrict subset of another linearly ordered subset.
+	A linearly \marginleft{Maximal subset} ordered subset (i.e. a chain) is \textbf{maximal} if it is not a strict subset of another linearly ordered subset.
 \end{defn}
@@
 \begin{prop}[Kuratowski's Lemma or Hausdorff maximal principle]
-	Every linearly ordered subset is the subset of some maximal linearly ordered subset.
+	Every chain is the subset of some maximal chain (i.e. maximal linearly ordered subset).
 \end{prop}
@@
 	A \textbf{discrete order} \marginleft{Discrete order} is an order in which every element that has a predecessor/successor has an immediate predecessor/successor.
+A \textbf{discrete linear order} is an order that is both linear and discrete.
+Be aware that discrete order often refers to an anti-chain.
 \end{defn}
-
-Note that discrete order may also mean an order in which every element is only related to itself (i.e. the partial order is an equivalence relation).
 
@@
 		\end{align*}
-		A complete lattice is \textbf{infinitely distributive} if the same holds over infinite operations.
+		A complete lattice is \textbf{infinitely distributive}, or \textbf{completely distributive} if the same holds over infinite operations.
 		\item[Complemented] A bounded lattice is \textbf{complemented} if every element has a complement. It is \textbf{uniquely complemented} if the complement is unique.
@@
 \begin{prop}
-	A lattice is distributive if an only if it does not contain an $M_3$ (diamond) or $N_5$ (pentagon) sublattice.
+	A lattice is distributive if and only if it does not contain an $M_3$ (diamond) or $N_5$ (pentagon) sublattice.
 \end{prop}
@@
 \begin{remark}
-	The completion is constructed by finding pairs $(A, B)$ of subsets $A, B \subseteq X$ such that $A$ is the set of lower bounds for $B$ and $B$ is the set of upper bounds of $A$.
+	The Dedekind-MacNeille completion is constructed by finding pairs $(A, B)$ of subsets $A, B \subseteq X$ such that $A$ is the set of lower bounds for $B$ and $B$ is the set of upper bounds of $A$.
 For the rationals $\mathbb{Q}$, these pairs correspond to Dedekind cuts, and the completion of $\mathbb{Q}$ as a lattice is the lattice of the reals $\mathbb{R}$.
@@
 		\item[Lattice] $(X, \vee, \wedge)$ is a lattice
-		\item[Distributivity] $a \vee (b \wedge c) = (a \vee b) \wedge (a \vee b)$
-and $a \wedge (b \vee c) = (a \wedge b) \vee (a \wedge b)$
+		\item[Distributivity] $a \vee (b \wedge c) = (a \vee b) \wedge (a \vee c)$
+and $a \wedge (b \vee c) = (a \wedge b) \vee (a \wedge c)$
 		\item[Identity] $a \vee \bot = a$
@@
 \begin{prop}[Stone's representation theorem]
-	Every distributive lattice is order isomorphic to a lattice of subsets ordered by inclusion.
+	Every bounded distributive lattice is order isomorphic to a lattice of subsets ordered by inclusion.
 \end{prop}

### 2025-09-22 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
NSF proposal

\begin{center}
	\textbf{Title} \\
	Christine A. Aidala, University of Michigan
\end{center}
%\vspace{-8mm}

\section{Introduction}

\section{Proposed research program}

\subsection{Intellectual merit}

\section{Summary of proposed research program}

\noindent
\textbf{Assessment of intellectual merit:} \\

\noindent
\textbf{Relationship to other current funding:} The PI's current funding from the NSF Division of Physics is for work on LHCb experiment at CERN, and from the NSF AccelNet program is to increase collaboration on quantum chromodynamics (QCD) research between the U.S.~and Latin America.  The PI's funding from the Department of Energy is for work at the Relativistic Heavy Ion Collider (RHIC) on the PHENIX and sPHENIX experiments.  These grants are for a completely separate line of research. Her grant from the John Templeton Foundation, titled ``Why classical probability and classical information theory are incompatible with quantum mechanics and quantum contextuality,'' supports the development of novel non-additive measures (set functions) suitable for quantifying the number of states in quantum mechanics~\cite{aop-nonaddmeas}.  There is no direct overlap with the activities proposed here.  \\

\noindent
\textbf{Relationship to other commitments of PI:} PI Aidala has been continuously involved in high-energy experimental QCD since 2001.  Her separate line of theoretical research on the mathematical foundations of physics is relatively new, with her first publication in 2018~\cite{aop-phys-blueprint}. She has successfully split her efforts between her different research lines for the past six years, with her work related to the Assumptions of Physics project producing 14 peer-reviewed publications, 1 more submitted paper, and a published technical book since 2018.  She is moreover in the process of reducing the number of personnel she supervises on QCD research in order to have more time for her theoretical work---four of her Ph.D.~students in QCD completed their degrees in the last 18 months, and two more are expected to graduate before the start of the proposed project.

\section{Personnel}
The proposed research will be performed by PI Christine Aidala, her collaborator Gabriele Carcassi, and Ph.D.~student Tobias Thrien, all part of the Physics Department at the University of Michigan.  See also the section on ``Qualifications of the PI and Research Team'' below.  \\
TODO: Add Alejandro assuming he agrees.

Aidala as PI will be responsible for coordinating all of the work, and she will serve as the academic advisor for Thrien.  She will identify suitable journals for publication and conferences at which to disseminate the results, and oversee the direction of the work.

Carcassi will serve as the technical lead.  He is (currently part-time) staff in the U-M Department of Physics, a Research Area Specialist Senior, with his salary fully supported through grants.  He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until Thrien is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.  With training in engineering and mathematics, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.

Thrien entered the U-M Physics Ph.D.~program in Fall 2024.  While Carcassi will be the primary individual responsible for ``outward-facing'' technical work, Thrien will focus primarily on inward-facing technical work.  \\
TODO: Be more specific about Thrien's work, per reviews from last year.

%\begin{figure*}[htb]
%	\centering
%	\includegraphics[scale=.6]{figures/PersonnelTimelineNSF2022.png}
%	\caption{Estimated timeline of research activities.}
%	\label{fig:personnelTimeline}
%\end{figure*}

\section{Broader impacts}
In addition to disseminating the goals and results of the project through traditional means such as peer-reviewed publications and presentations at workshops and conferences, the goals and results of the project will be disseminated to wider audiences by leveraging existing social media and other outreach infrastructure that has already been developed for the broader Assumptions of Physics project, in particular, Carcassi's two YouTube channels, as described below in the section on qualifications of the research team.  One of them is aimed at the level of individuals with upper-level undergraduate training in physics and/or mathematics, and the second, more technical channel is aimed at professional researchers, for which he creates videos presenting detailed technical results or describing open technical questions, leveraging the expertise of others to help advance research goals.  Graduate student Tobias Thrien originally discovered the AoP project through Carcassi's YouTube channel while a Bachelor's student in Germany several years ago.  He has since completed a Master's in theoretical cosmology in Germany and is currently a first-year Ph.D.~student at U-M, having come to Michigan specifically to work on the AoP project with the PI and Carcassi.  Regarding the impact of posting open technical problems, our first collaborative paper that resulted from someone reaching out to us after seeing a video, on classical mechanics as the high-entropy limit of quantum mechanics, is currently under review at a journal~\cite{aop-classicallimit}.  We were previously completely unaware of the researcher and his work at the University of Innsbruck, Austria. Carcassi additionally writes monthly blog-style essays related to the AoP project and advertised through social media platforms such as Facebook and LinkedIn.  Thus, the results of the proposed research will reach a much wider audience than through the publications and conference talks alone.  Further dissemination of the open-research approach may inspire researchers in completely different areas to adopt a similar approach for their own projects, with potential acceleration of research progress in particular within interdisciplinary areas, where not all necessary expertise may be available within a single research community.

TODO: Paragraph about impact of unifying disparate branches of physics, and eventual impact on physics education.\\

\noindent
\textbf{Assessment of broader impacts:}  The reach and impact of videos related to the proposed work will be assessed by the number of views they receive and by other measures of audience engagement such as comments posted or subsequent email contacts.  The impact of unifying disparate branches of physics will need to be assessed on a longer-term timescale than that of the grant, but early indicators could include talk invitations and paper citations. \\
TODO: Consider ways to assess engagement in more detail, per last year's reviews.

\section{Qualifications of the PI and research team}
PI Christine Aidala is a highly recognized leader within physics, having received an NSF CAREER Award, a Sloan Fellowship, and a Presidential Early Career Award for Scientists and Engineers. She was a 2019-20 U.S.~Fulbright Scholar and is a Fellow of the American Physical Society (APS).  She served on the National Academy of Sciences Committee on a U.S.-Based Electron-Ion Collider Science Assessment 2016-18 and is currently serving on the U.S.~Nuclear Science Advisory Committee.  She served as an elected member of the Executive Committee of the APS Division of Nuclear Physics (DNP) 2021-23, and since 2021 she has served as a DNP Ally to ensure that DNP meetings are welcoming and inclusive environments.  Much of her work has been in quantum chromodynamics; however, in 2015 she initiated the Assumptions of Physics project with her collaborator and husband Gabriele Carcassi, and since 2018 they have so far published together 14 peer-reviewed articles~\cite{aop-phys-blueprint,aop-topExpDisting,aop-HamQuantInfo,Carcassi:2021,aop-spacetimeStruct,aop-HamConsInfoEnt,Carcassi2021four,Carcassi:2022bpm,aop-HamPriv,aop-commonLogical,aop-action,aop-nogo,aop-nonaddmeas,aop-unphysHilbert} as well as a technical book on the mathematical foundations of physics~\cite{aop-book}, with one additional article currently under review and publicly available as a preprint~\cite{aop-classicallimit}.  She is the PI on a grant from the John Templeton Foundation that supports the development of novel measure theoretic structures for quantum mechanics. \\

TODO: Add info for Alejandro.

Carcassi is the technical lead for the Assumptions of Physics project.  He is the Co-PI on the Templeton Foundation grant with Aidala.  Related to the AoP project, he has given seminars to audiences of physicists, mathematicians, and philosophers at U-M, Harvard, Caltech, the University of Pavia, Italy, the Munich Center for Mathematical Philosophy, Germany, and the Wolfram Institute, and he and Aidala gave a joint public talk for the Michigan State University Advanced Studies Gateway.  His training in engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from theoretical computer science, information theory, systems theory, signal processing, and operations research.  His position as U-M Physics Department staff without teaching responsibilities provides him with time and flexibility to dedicate to research related to the AoP project such as the work proposed here. \\
TODO: Add Kansas, Saturday Morning Physics, invited talk at AMS meeting, also earlier Topology conference talk?

Together, Aidala and Carcassi are uniquely positioned to advance the mathematical foundations of physics, with Aidala's traditional training in physics and traditional academic position complementing Carcassi's formal training in engineering and his university staff position, which for example gave him the time and flexibility to audit multiple graduate Mathematics courses at U-M.  For the AoP project, their published work so far has covered a wide range of topics.  The ``physical mathematics'' approach within their work is an approach to the mathematical foundations of physics that seeks to construct mathematical structures strictly from axioms and definitions that can be rigorously justified from physical requirements~\cite{aop-book}.  For example, they have shown that experimental distinguishability is mathematically captured by Hausdorff topologies~\cite{aop-book,aop-topExpDisting}, and they have argued that at the smallest space-time scales, at which geometric information becomes in principle experimentally inaccessible, there should still in principle be a regime in which topological structure remains experimentally accessible~\cite{aop-spacetimeStruct}.  Their ``reverse physics'' approach instead is an approach to the foundations of physics that analyzes known theories to identify those physical principles and assumptions that can be taken as their conceptual foundation~\cite{aop-book,Carcassi:2022bpm}. Their reverse physics approach has led for example to a geometric and physical interpretation of the principle of least action~\cite{aop-action} and to a demonstration that the uncertainty principle in quantum mechanics emerges solely from the lower limit on the entropy~\cite{Carcassi:2022bpm}. In other work, they have for example reduced the number of postulates of quantum mechanics, proving that the tensor product postulate for composite systems can be derived from the state and measurement postulates~\cite{Carcassi2021four}, and they have offered a clearer characterization of Shannon information entropy~\cite{Carcassi:2021}.

Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds, and in 2025 the U-M Rackham Distinguished Graduate Mentor Award.  She has served as the dissertation chair for 10 graduate students in Physics and Applied Physics who have already completed their doctorates, as co-chair for one in Applied and Indisciplinary Mathematics and another in Applied Physics, and is currently the dissertation chair for an additional 4 Ph.D.~students in Physics.  Aidala and Carcassi have also successfully published with former U-M Ph.D.~students in Mathematics~\cite{aop-spacetimeStruct} and Philosophy~\cite{aop-HamPriv} whose primary dissertation work was with other faculty members.  In 2020 Aidala received the U-M John Dewey Award for long-term commitment to the education of undergraduate students.  Between Aidala and Carcassi, they have mentored more than 50 undergraduate students in research since coming to U-M in 2012.  \\

TODO: Add info on summer schools here?

TODO: Add info for Alejandro.

Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with more than 17,500 subscribers currently.  He posts a new video monthly, generally targeted at individuals with an advanced undergraduate background in physics and/or mathematics.  Collectively, his videos on this channel have received over 350k views [TODO: update]. He additionally has a second, more technical YouTube channel (@AssumptionsofPhysicsResearch) aimed at professional researchers, with more than 1600 subscribers.  He uses this channel as a venue to disseminate research results at a more technical level, including video versions of our preprints or published papers as well as conference presentations.  He moreover uses the technical channel as an innovative approach to try to ``crowd-source'' technical expertise to advance the research more efficiently.  Given the breadth of the broader AoP project, there is no single community with the appropriate technical background across all relevant and potentially relevant areas of mathematics and physics.  So far this ``crowd-sourcing'' approach has led to multiple instances of viewers with whom we have had no previous contact pointing us to valuable reference materials, and it recently led to our first coauthored paper with a researcher who reached out to us after seeing an open problem video, on ``Classical mechanics as the high-entropy limit of quantum mechanics''~\cite{aop-classicallimit}. We were previously completely unaware of the researcher and his work in Innsbruck, Austria.  In other forms of reach beyond academia, Carcassi was interviewed in 2024 by Curt Jaimungal for the ``Theories of Everything'' podcast~\cite{Carcassi-ToEInterview}, which has 500k subscribers.  Carcassi additionally writes monthly blog-style essays related to the AoP project and advertised through social media platforms such as Facebook and LinkedIn.  In more creative outreach work, in April 2024 at U-M he performed an original musical comedy lecture, ``An Engineer in the Foundations of Physics,'' which drew approximately 100 attendees.

Graduate student Tobias Thrien, for whom partial support is requested, originally discovered the Assumptions of Physics project through Carcassi's YouTube channel while a Bachelor's student in Germany several years ago.  He has since completed a Master's in theoretical cosmology in Germany and is currently a second-year Ph.D.~student at U-M.  He attended Aidala and Carcassi's inaugural online summer school on the AoP project in June 2024 and is already highly familiar with the broader project, its goals, and its approaches.

\section{Results from prior NSF support}
\subsection{Intellectual merit of prior NSF support}

Aidala has been a successful steward of NSF funds since receiving a CAREER Award in 2015.  Her prior research awards have been through the Physics Division, Nuclear Physics - Experiment program: \emph{CAREER: Valence and sea quark dynamics at Fermilab}, 2015-21, and \emph{Studying quantum chromodynamics at LHCb}, 2020-present (renewed in 2023).  She is additionally a Co-PI on an NSF AccelNet award, \emph{AccelNet-Design: Inter-American Network of Networks on Quantum Chromodynamics Challenges}, 2021-present, aimed at further developing collaboration in quantum chromodynamics research between the U.S.~and Latin America.  Her QCD research has focused on advancing the multidimensional quark-gluon structure of the proton and on studying the processes by which scattered quarks or gluons form new bound states (hadrons).  Her group's NSF-funded work includes results published in Nature~\cite{SeaQuest:2021zxb}, Physical Review Letters~\cite{LHCb:2022tbc,LHCb:2019qoc}, Physical Review D Letters~\cite{LHCb:2022rky}, and Physical Review C~\cite{SeaQuest:2022xdu}, among other journals.  Regarding the specific scientific advancements of some of the work, Aidala and her collaborators have studied antimatter in the proton and found surprising evidence that there is an excess of down ``flavored'' antiquarks relative to up flavored antiquarks in the proton for antiquarks carrying as much as 45\% of the proton's total momentum~\cite{SeaQuest:2021zxb,SeaQuest:2022xdu}.  Her group has also led the first-ever measurements of identified hadron species production by high-energy light-mass quarks in proton-proton collisions, multidifferential in multiple kinematic variables simultaneously, showing that existing models poorly predict the relative production rates of some species~\cite{LHCb:2022rky}. [TODO: Add dead cone effect] This proposal is the second time she is submitting to the Applied Mathematics program, based on the research direction in the mathematical foundations of physics that she initiated more recently, which has so far produced 14 peer-reviewed papers since 2018 as well as a technical book, as described above in the section on qualifications of the PI and research team.  \\

TODO: Prior NSF support for Alejandro?

\subsection{Broader impacts of prior NSF support}
Aidala's group has been leading the open data efforts of the LHCb experiment at the Large Hadron Collider at CERN.  Of particular note, they have developed a user-friendly web-based application to securely access and provide analysis documentation for large-scale data sets in high-energy physics, the ``NTuple Wizard''~\cite{Aidala:2023dai}.  It is currently in beta-release to provide access to open data sets from LHCb, and individuals from other experimental collaborations have expressed interest in potentially adapting it to create a similar application to facilitate access to their own open data sets and corresponding documentation.

[TODO: Update and rewrite to avoid DEI language.] In addition, a variety of activities under Aidala's prior NSF grants have contributed directly to training junior scientists and to diversity in the STEM workforce.  A woman postdoctoral scholar was involved 2019-22. Including students with support from on and off of the grant, 8 Ph.D. students and 4 undergrads have been involved in the project since July 2020. An additional 3 distinct Ph.D. students and 7 undergraduates worked on the projects supported by Aidala's CAREER grant 2015-20.  Of the grad students, 2 have been African-American men, 2 Caucasian men, 2 Latino men, 1 an Arabic man, and 4 women, of which 2 Latina.  Former African-American grad student Bryan Ramson is now a permanent staff scientist at Fermilab; former Latina grad student Catherine Ayuso is now Head of Special Projects and Research Manager at Entanglement, Inc.  The others are in postdoctoral positions or still students in the group.  The undergrads included 1 Latino, 1 man who was openly autistic, and 5 women from Mt.~Holyoke College, among which were international students from the Philippines, Bangladesh, and China. Four of the grad/undergrad students openly identify as LGBT.  All Mt.~Holyoke students have continued in STEM.  One received an M.S. from Stanford in Statistics and now works for Amazon, the other four are now Ph.D. students at Harvard, Michigan, Virginia Tech, and Vanderbilt in Applied Physics, Physics, or Electrical Engineering.  \\

TODO: Prior NSF support for Alejandro?

\newpage

### 2025-09-22 Proposals/2025-NSF-EntropicGeometry/ProjectSummary.tex
NSF proposal

\thispagestyle{empty}
\noindent
\textbf{Overview:}\\

\noindent
\textbf{Intellectual Merit:}  \\

\noindent
\textbf{Broader impacts:}  \\

The proposed activities additionally include significant outreach through social media, primarily through YouTube.  Videos will be produced to disseminate the results of the research, with some videos aimed at the level of an advanced undergraduate in mathematics and/or physics, furthering education.  Other videos will be aimed at active researchers in various fields of mathematics and physics, providing a new mathematical framework as well as a new perspective linking traditionally disparate realms of physics that may impact tools or thinking in their own research.  In the spirit of an open-research approach that moreover leverages ``crowd'' contributions, YouTube videos aimed at researchers will additionally be used to advertise open problems that the project personnel may not be able to expediently resolve themselves. Further dissemination of the open-research approach may inspire researchers in completely different areas to adopt a similar approach for their own projects, with potential acceleration of research progress in particular within interdisciplinary areas, where not all necessary expertise may be available within a single research community.

### 2025-09-22 BareMinima/OrderTheory/BareMinimumOrderTheory.tex
Reviewed abstract for bare mininum

@@
 \begin{abstract}
-A summary overview of order theory, which can be used as a reference, refresher or to be aware of what one does not know.
+	A condensed overview of order theory.
+Bare minima are meant to give a rough overview, by no means complete, of the subject so that one at least knows what there is to know.
+It is mainly intended as background for those interested in participating in the Assumptions of Physics (\url{https://assumptionsofphysics.org}) project.
 \end{abstract}

### 2025-09-22 BareMinima/SetTheory/BareMinimumSetTheory.tex
SetTheory: moving families at the end of the first section
Reviewed abstract for bare mininum

@@
 \begin{abstract}
-This note presents a condensed summary of set theory which can function as a crash course, refresher and/or reference.
-Bare minima are meant to give a rough overview, by no means complete, of the subject to the intellectually curious, particularly in the context of foundational questions in physics.
-This work is part of Assumptions of Physics (\url{https://assumptionsofphysics.org}), a project that aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+A condensed overview of set theory.
+Bare minima are meant to give a rough overview, by no means complete, of the subject so that one at least knows what there is to know.
+It is mainly intended as background for those interested in participating in the Assumptions of Physics (\url{https://assumptionsofphysics.org}) project.
 \end{abstract}
@@
 
-\subsection{Indexed families and sequences}
-
-\begin{defn}[Families]
-	An \textbf{indexed family}, \marginleft{Families, sequences: $\{x_i\}_{i \in I}$, $\{x_i\}_{i=1}^{\infty}$} or simply a \textbf{family}, $\{x_i\}_{i \in I}$ is a collection of elements identified by an \textbf{index set} $I$.
-More formally, an indexed family is a tuple $( X, I, x )$ where $X$ and $I$ are sets and $x: I \to X$ is a function such that $i \mapsto x_i$.
-A \textbf{sequence} is a family for which the index set is a contiguous interval of natural numbers.
-A \textbf{finite sequence} of $n$ elements is noted $\{x_i\}_{i=1}^{n}$ while an \textbf{infinite sequence} is noted $\{x_i\}_{i=1}^{\infty}$.
-\end{defn}
-
-\begin{defn}[Operations on families of sets]
-	Let $\{A_i\}_{i \in I}$ \marginleft{$\bigcap\limits_{i \in I} A_i $, $\bigcup\limits_{i \in I} A_i$} be a family of sets, the operations of intersection $\bigcap\limits_{i \in I} A_i $ and union $\bigcup\limits_{i \in I} A_i$ can be extended to span all members.
-\end{defn}
-
-\begin{defn}[Set-theoretic limit]
-	Let $\{A_i\}_{i=1}^{\infty}$ \marginleft{Set limit: \\ $\lim\limits_{i \to \infty} A_i = A$} be a sequence of sets.
-We define the \textbf{limit infimum} as $\liminf\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1}\bigcap\limits_{j \geq i} A_j$ and the \textbf{limit supremum} as $\limsup\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1}\bigcup\limits_{j \geq i} A_j$.
-If the two are equal to the same set then we say the sequence \textbf{converges}, the \textbf{set-theoretic limit} $\lim\limits_{i \to \infty} A_i = A$ exists and it is equal to that set.
-\end{defn}
-
-\marginright{~\\
-	\def\setA{ (-1,0) circle (2) }
-	\def\setB{ (1,0) circle (2) }
-	\begin{tikzpicture}[scale = 0.3]
-		\draw (0,0) ellipse (5 and 4); 
-		\draw (0.5,-0.7) ellipse (3.5 and 2.7);
-		\draw (1.05,-1.3) ellipse (2.2 and 1.6);
-		\draw (1.25,-1.45) ellipse (1.6 and 1.05);
-		\draw (1.3,-1.6) ellipse (1.05 and 0.55);
-		\node at (0,-5) {$A_i \supset A_{i+1}$};
-		\node at (-2.6, 2.1) {$A_1$};
-		\node at (-1.05,0.55) {$A_2$};
-		\node at (1.3,-1.6) {$...$};
-	\end{tikzpicture} 
-}
-\begin{defn}[Monotone sequence]
-	A \textbf{monotone sequence} \marginleft{Monotone sequence: \\ $A_1 \subseteq A_2 \subseteq ...$\\ $A_1 \supseteq A_2 \supseteq ...$} is a sequence of sets $\{A_i\}_{i=1}^{\infty}$ where each set is a subset (or superset) of the preceding.
-More specifically, we distinguish four cases:
-	\begin{description}
-		\item[Increasing] when $A_i \subseteq A_{i+1}$ for all $i \geq 1$
-		\item[Decreasing] when $A_i \supseteq A_{i+1}$ for all $i \geq 1$
-		\item[Strictly increasing] when $A_i \subset A_{i+1}$ for all $i \geq 1$
-		\item[Strictly decreasing] when $A_i \supset A_{i+1}$ for all $i \geq 1$
-	\end{description}
-\end{defn}
-
-\begin{prop}[Monotone convergence]
-	All monotone \marginleft{Monotone convergence} sequences converge.
-For increasing sequences the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1} A_i$, while for decreasing sequences  the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1} A_i$.
-\end{prop}
-
 \subsection{Homogeneous relations, equivalences and orders}
@@
 
+\subsection{Indexed families and sequences}
+
+\begin{defn}[Families]
+	An \textbf{indexed family}, \marginleft{Families, sequences: $\{x_i\}_{i \in I}$, $\{x_i\}_{i=1}^{\infty}$} or simply a \textbf{family}, $\{x_i\}_{i \in I}$ is a collection of elements identified by an \textbf{index set} $I$.
+More formally, an indexed family is a tuple $( X, I, x )$ where $X$ and $I$ are sets and $x: I \to X$ is a function such that $i \mapsto x_i$.
+A \textbf{sequence} is a family for which the index set is a contiguous interval of natural numbers.
+A \textbf{finite sequence} of $n$ elements is noted $\{x_i\}_{i=1}^{n}$ while an \textbf{infinite sequence} is noted $\{x_i\}_{i=1}^{\infty}$.
+\end{defn}
+
+\begin{defn}[Operations on families of sets]
+	Let $\{A_i\}_{i \in I}$ \marginleft{$\bigcap\limits_{i \in I} A_i $, $\bigcup\limits_{i \in I} A_i$} be a family of sets, the operations of intersection $\bigcap\limits_{i \in I} A_i $ and union $\bigcup\limits_{i \in I} A_i$ can be extended to span all members.
+\end{defn}
+
+\begin{defn}[Set-theoretic limit]
+	Let $\{A_i\}_{i=1}^{\infty}$ \marginleft{Set limit: \\ $\lim\limits_{i \to \infty} A_i = A$} be a sequence of sets.
+We define the \textbf{limit infimum} as $\liminf\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1}\bigcap\limits_{j \geq i} A_j$ and the \textbf{limit supremum} as $\limsup\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1}\bigcup\limits_{j \geq i} A_j$.
+If the two are equal to the same set then we say the sequence \textbf{converges}, the \textbf{set-theoretic limit} $\lim\limits_{i \to \infty} A_i = A$ exists and it is equal to that set.
+\end{defn}
+
+\marginright{~\\
+	\def\setA{ (-1,0) circle (2) }
+	\def\setB{ (1,0) circle (2) }
+	\begin{tikzpicture}[scale = 0.3]
+		\draw (0,0) ellipse (5 and 4); 
+		\draw (0.5,-0.7) ellipse (3.5 and 2.7);
+		\draw (1.05,-1.3) ellipse (2.2 and 1.6);
+		\draw (1.25,-1.45) ellipse (1.6 and 1.05);
+		\draw (1.3,-1.6) ellipse (1.05 and 0.55);
+		\node at (0,-5) {$A_i \supset A_{i+1}$};
+		\node at (-2.6, 2.1) {$A_1$};
+		\node at (-1.05,0.55) {$A_2$};
+		\node at (1.3,-1.6) {$...$};
+	\end{tikzpicture} 
+}
+\begin{defn}[Monotone sequence]
+	A \textbf{monotone sequence} \marginleft{Monotone sequence: \\ $A_1 \subseteq A_2 \subseteq ...$\\ $A_1 \supseteq A_2 \supseteq ...$} is a sequence of sets $\{A_i\}_{i=1}^{\infty}$ where each set is a subset (or superset) of the preceding.
+More specifically, we distinguish four cases:
+	\begin{description}
+		\item[Increasing] when $A_i \subseteq A_{i+1}$ for all $i \geq 1$
+		\item[Decreasing] when $A_i \supseteq A_{i+1}$ for all $i \geq 1$
+		\item[Strictly increasing] when $A_i \subset A_{i+1}$ for all $i \geq 1$
+		\item[Strictly decreasing] when $A_i \supset A_{i+1}$ for all $i \geq 1$
+	\end{description}
+\end{defn}
+
+\begin{prop}[Monotone convergence]
+	All monotone \marginleft{Monotone convergence} sequences converge.
+For increasing sequences the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcup\limits_{i \geq 1} A_i$, while for decreasing sequences  the limit is given by $\lim\limits_{i \to \infty} A_i = \bigcap\limits_{i \geq 1} A_i$.
+\end{prop}
+
 \section{Axiomatic set theory}

### 2025-09-23 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Add title and notes from last week

@@
 \begin{center}
-	\textbf{Title} \\
+	\textbf{A generalized axiomatic basis for entropic structures in physical theories} \\
 	Christine A. Aidala, University of Michigan
@@
 \section{Introduction}
+
+Notes:
+
+Generalized ensemble spaces + geometry defined from entropy; full set of axioms
+
+-Develop axiomatics for the entropy such that they’re physically motivated.
+Not taking starting points because they’re convenient mathematically, nor because we think one physical theory (e.g.
+quantum mechanics) is the foundational one.
+
+-See if this type of framework will help constrain the problem in infinite dimensions (motivation, not goal).
+Need to generalize the Poisson bracket structure in infinite dimensions—relationship between entropy and Poisson brackets used as Lie brackets.
+Well-known open research problem that there’s no general theory for infinite-dimensional analysis/infinite-dimensional Lie groups.
+
+Motivation (but not goal): Eventual framework to generalize classical and quantum information theory
+
+1. To answer fundamental questions.
+
+2. To do classical and quantum info in a more generalized setting
+
+3. Give a structure for infinite-dimensional analysis that’s constrained enough that mathematicians can prove theorems that are useful for physics
+
+Have: 1. Topological structure; 2. Convex structure; 3. Geometric structure.
+
+Open problems: 1. Entropic open balls can generate topology? 2. Are all axioms independent?   …
+
+To determine which fundamental structures/properties are needed for all physical theories.
+
+Open with: One of the difficulties with mathematical physics is that physicists aren’t good at specifying the mathematical requirements such that mathematicians can work effectively on the problems.
+
+Mention economists evidently aware of topological issues more than physicists.
 

### 2025-09-30 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
2025NSF: working on the proposal description

@@
 \section{Introduction}
+
+Physics is currently a patchwork of different theories with no common underlying foundation.
+As part of our larger project, Assumptions of Physics, we aim to develop an axiomatic basis for states and processes that can be equally applied to different theories, such as classical and quantum mechanics.
+Different theory are then recovered as specializations of the more general structure.
+
+A core insight that is guiding this enterprise is that geometric structure in both classical and quantum mechanics are equivalent to entropic structures.
+In classical mechanics, a single degree of freedom is a two dimensional symplectic manifold.
+In this case, symplectic form $\omega$ is a volume form and therefore specifying $\omega$ is equivalent to specifying the Liouville measure $\mu(U) = \int_U \omega$.
+In classical statistical mechanics, the entropy of a statistical ensemble is given by $S(\rho) = - \int \rho \log \rho d\mu$ where $\rho$ is the probability density associated to the ensemble.
+For a uniform distribution with support $U$, the entropy reduce to $S(\rho_U) = \log \mu(U)$.
+Since the logarithm is an invertible function, one can recover the measure from the entropy.
+That is, specifying the entropy associated to all uniform distributions allows us to reconstruct the Liouville measure, and therefore the symplectic form.
+In the general case, Darboux coordinates allow, at least locally, to factorize the space into multiple independent degrees of freedom, where the volume is the product of the corresponding areas.
+The corresponding entropic property is that the entropy for the joint distribution is the sum of the marginals if they are independently distributed.
+Therefore, if the distribution of all possible marginals are specified, the symplectic form can be recovered.
+
+In quantum mechanics, the geometry is given by the Born rule
+\begin{equation}
+	p(\phi|\psi) = \frac{\< \phi | \psi \>\< \psi | \phi \>}{\< \psi | \psi \>\< \phi | \phi \>}.
+\end{equation}
+Given two states $\psi$ and $\phi$, the ensemble $\rho_{\psi\phi}$ given by their equal mixture has entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi,\psi)}}{2},\frac{1 - \sqrt{p(\phi,\psi)}}{2} \right)$ where $I(x,y) = -x \log x -y \log y$ is the Shannon entropy.
+The entropy as a function of $p(\psi,\phi)$ is invertible.
+Therefore, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the state space.
+
+While the states spaces of classical (i.e.
+symplectic manifolds) and quantum mechanics (i.e.
+projective Hilbert spaces) are rather different, the space of ensembles of both have a similar structure.
+They are both convex subsets of a topological vector space, equipped with a strictly convex entropy function.
+The core idea, then, is to use ensembles as a way to generalize the notion of physical states, and the entropy to characterize the geometry in a way that is independent of the particulars of each physical theory.
+In the past year we have done significant preliminary work that shows that not only this approach is possible, but is extremely productive.
+The proposed work is to turn those preliminary insight in an axiomatic specification that is fully physically motivated.
+Such an approach, if successful, would at least provide the following benefits:
+\begin{enumerate}
+	\item clarify exactly what physical premises guarantee the basic geometric structures of physical theories, allowing a more principled formulation of current physical theories and search for new ones;
+	\item provide a generalized structure for information theory, possibly generalizing results from both classical and quantum information theory
+	\item provide useful constrains to more precisely formulate physical theories in the infinite dimensional case.
+\end{enumerate}
+
+\section{Proposed research program}
+
+In a nutshell, the goal is to find a very limited set of axioms that can be guaranteed to be valid in any physical theory, and that allow us to generalize a good number of results from all physical theories.
+The notion of statistical ensemble (i.e.
+the collection of all outputs of a preparation procedure) is particularly suitable for generalization for a number of both practical and conceptual considerations.
+Our preliminary work shows that there are three, possible four, distinct set of requirements for an ensemble space $\Ens$, the set of ensemble given by a physical theory.
+
+The first requirement is that of experimental verifiability of the theory.
+This has been thoroughly investigated in our previous work that established a precise correspondence between experimental verifiable statements (i.e.
+statement associated with an experimental test that terminates successfully in finite time if and only if the statement is true) and the open set of a ``natural'' topology, and between theoretical statements (i.e.
+statement associated with an experimental test without termination guarantees) and Borel sets.
+Real and integer numbers, for example, are recovered from an idealized model representing references of a reference system.
+The key insight is that the set theoretic operations represent a two-valued logic system (i.e.
+truth or falsehood of statements) while the topology keeps track of a three-valued logic system (i.e.
+interior, exterior and boundary represent respectively successful termination, unsuccessful termination and non-termination not terminate).
+This work gives us a solid foundation on which to build, and guarantees that an ensemble space $\Ens$ is, at least, a $\mathsf{T}_0$ second countable topological space.
+
+The second requirement is the ability to perform statistical mixtures.
+Given two ensembles $\ens[a], \ens[b] \in \Ens$, we can always choose a probability $p \in [0,1]$ and select $\ens[a]$ $p$ percent of the times and $\ens[b]$ $\bar{p} = 1-p$ percent of the times.
+This gives the ensemble space a convex structure as $p \ens[a] + \bar{p} \ens[b] \in \Ens$ for all $\ens[a], \ens[b] \in \Ens$ and $p \in [0,1]$.
+This is insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs).
+The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to physically untenable cases).
+Rather, we let the topology do the work.
+That is, infinite convex combinations exist only if they are the topological limit of finite convex combinations.
+On physical grounds, the mixing operation must be continuous in the natural topology defined by experimental verifiability, much like addition and scalar multiplication are continuous in a topological vector space, leading to the idea of a topological convex space.
+Unfortunately, we haven't found such a structure already studied in the literature.
+While it would be extremely useful for us to study these spaces on their own merit, we choose to concentrate on problems where our expertise in the mathematical foundations of physical theories is more critical, as in the next requirement.
+
+The third requirement is characterizing the variability of elements within a statistical ensemble.
+Though the instance within ensembles are never formalized, the variability within an ensemble is physically significant as it will influence, for example, the variability of statistical measurements.
+As we showed in previous work, entropy quantifies the variability of the elements within a distribution, and which gives a physically meaningful starting point.
+The proposed work is to find the exact specification of the axiom to see whether it can also guarantee that ensemble spaces must be metric spaces.
+
+The hardest conceptual problem is that entropy is defined differently in each space.
+In a classical discrete space, the entropy is $-\sum p_i \log p_i$; in classical statistical mechanics is $- \int \rho \log \rho d\mu$ where $\mu$ is the Liouville measure; in quantum mechanics is $-\tr(\rho \log \rho)$.
+How can we give a single definition for these very different expressions?
+Note that the expressions are different precisely because they are in terms of the specific representation of ensembles in specific theory.
+Our approach is to simply characterize the entropy by its properties with respect to the topology and the convex structure.
+In our preliminary properties, for example, the entropy $S : \Ens \to \mathbb{R}$ has the following properties:
+\begin{itemize}
+	\item Continuity
+	\item Strict concavity: $S(p\ens[a] + \bar{p} \ens[b]) \geq p S(\ens[a]) + \bar{p} S(\ens[b])$ with the equality holding if and only if $\ens[a] = \ens[b]$
+	\item Upper variability bound: there exists a universal function $I(p_1, p_2)$ (i.e.
+the same for all ensemble spaces) such that $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p, \bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])$; if the equality holds, $\ens[a]$ and $\ens[b]$ are orthogonal, noted $\ens[a] \ortho \ens[b]$
+	\item Mixtures preserve orthogonality: $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$
+\end{itemize}
+These properties are justified by the physics.
+For example, if we are performing a statistical mixture between two ensemble, the average variability cannot decrease, which justifies the concavity.
+These relatively simple properties allow us, for example, to show that $I(p,\bar{p})$, the maximum entropy increase during mixing, must be $ - \kappa \left(p \log p + \bar{p} \log \bar{p}\right)$, the Shannon entropy.
+The proof is general, meaning that if we are right that those axioms must hold for any physical theory, then the Shannon entropy is of truly general applicability.
+Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is valid with an equal.
+That allows to recover the expression for the entropy in the three different cases.
+
+Our preliminary work shows that an entropy satisfying those axioms forces the convex space to be cancellative, meaning that it embeds into a vector space.
+That is, if we pick an internal point $\ens \in \Ens$, we can define a vector space $V_{\ens}$ such that $\Ens \subset V_{\ens}$.
+If we pick to different internal points $\ens \in \Ens$, we can show that there is an invertible affine map between $V_{\ens[a]}$ is $V_{\ens[b]}$.
+That is, the ensemble space embeds in an affine space.
+We can also show that it is ``directionally bounded,'' meaning its intersection with any affine line is finite.
+It is an open question whether the axioms are enough to prove that the embedding is continuous in a topological vector space, and whether the set is bounded.
+
+We can also show that the entropy can be used to define geometric structures directly in the ensemble space.
+For example, we can define the mixing entropy as $MS(\ens[a], \ens[b]) = S\left(\frac{1}{2}\ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2} S(\ens[a]) + \frac{1}{2} S(\ens[b])\right)$.
+If we express the entropy in bits, the mixing entropy satisfies the following:
+\begin{enumerate}
+	\item non-negativity: $MS(\ens[a], \ens[b]) \geq 0$
+	\item identity of indiscernibles: $MS(\ens[a], \ens[b]) = 0 \iff \ens[a]=\ens[b]$
+	\item unit boundedness: $MS(\ens[a], \ens[b]) \leq 1$
+	\item maximality of orthogonals: $MS(\ens[a], \ens[b]) = 1 \iff \ens[a] \ortho \ens[b]$
+	\item symmetry: $MS(\ens[a], \ens[b]) = MS(\ens[b], \ens[a])$
+\end{enumerate}
+It is almost a distance function, as it does not satisfy the triangle inequality, but it recovers the Jensen-Shannon divergence (JSD) in both classical and quantum mechanics.
+The JSD is the square of a distance function in both classical and quantum information theory, though our current axioms do not guarantee $MS$ to be so.
+The $MS$ can be used to create open balls, and show that the topology must be Hausdorff.
+The current axioms do not guarantee the open balls to be convex.
+
+If we assume we are in finite dimensions and that the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite which allows us to define a metric tensor.
+Let $V_{\ens[a]}$ be the embedding vector space with $\ens[a] \in \Ens$ as origin.
+This will be a differentiable manifold.
+On a chart whose coordinates are linear with respect to the affine structure we can define $g_{\ens}( v, w ) = - \frac{\partial^2 S}{\partial \ens ^2}( v, w )$ where $v$ and $w$ are elements of the tangent space at $\ens \in \Ens \subset V_{\ens[a]}$.
+Since all Hessian are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor.
+In classical and quantum mechanics it recovers the Fisher information metric.
+Note that, if were to show continuous embedding into a locally convex topological vector space, this construction can be extended to the infinite dimensional case.
+
+understand whether the existence of a physically meaningful entropy function is enough to guarantee that the space of statistical ensembles of any physical theory is a convex subset of a locally convex topological vector space equipped with a metric.
+That is, we should not simply ``declare'' it to be so, but make sure that the mathematical starting points are justified by physical arguments, so that we can
 

### 2025-10-03 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
2025NSF: draft of proposed work

@@
 
-A core insight that is guiding this enterprise is that geometric structure in both classical and quantum mechanics are equivalent to entropic structures.
+A core insight that is guiding this enterprise is that geometric structures in both classical and quantum mechanics are equivalent to entropic structures.
 In classical mechanics, a single degree of freedom is a two dimensional symplectic manifold.
@@
 In the general case, Darboux coordinates allow, at least locally, to factorize the space into multiple independent degrees of freedom, where the volume is the product of the corresponding areas.
-The corresponding entropic property is that the entropy for the joint distribution is the sum of the marginals if they are independently distributed.
+The corresponding entropic property is that the entropy for the joint distribution is the sum of the marginals if the independent degrees of freedom are independently distributed.
 Therefore, if the distribution of all possible marginals are specified, the symplectic form can be recovered.
@@
 \end{equation}
-Given two states $\psi$ and $\phi$, the ensemble $\rho_{\psi\phi}$ given by their equal mixture has entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi,\psi)}}{2},\frac{1 - \sqrt{p(\phi,\psi)}}{2} \right)$ where $I(x,y) = -x \log x -y \log y$ is the Shannon entropy.
-The entropy as a function of $p(\psi,\phi)$ is invertible.
-Therefore, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the state space.
+Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, mathematically $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi,\psi)}}{2},\frac{1 - \sqrt{p(\phi,\psi)}}{2} \right)$ where $I(x,y) = -x \log x -y \log y$ is the Shannon entropy.
+Therefore the entropy is an invertible function of $p(\psi,\phi)$.
+This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the state space.
 
@@
 The proposed work is to turn those preliminary insight in an axiomatic specification that is fully physically motivated.
-Such an approach, if successful, would at least provide the following benefits:
+If successful, such an approach will provide the following benefits:
 \begin{enumerate}
-	\item clarify exactly what physical premises guarantee the basic geometric structures of physical theories, allowing a more principled formulation of current physical theories and search for new ones;
-	\item provide a generalized structure for information theory, possibly generalizing results from both classical and quantum information theory
+	\item clarify exactly what physical premises guarantee the basic geometric structures of physical theories, allowing a more principled formulation of current physical theories as well as the search for new ones;
+	\item provide a generalized structure for information theory, possibly allowing future generalization of results from both classical and quantum information theory
 	\item provide useful constrains to more precisely formulate physical theories in the infinite dimensional case.
@@
 
-In a nutshell, the goal is to find a very limited set of axioms that can be guaranteed to be valid in any physical theory, and that allow us to generalize a good number of results from all physical theories.
-The notion of statistical ensemble (i.e.
-the collection of all outputs of a preparation procedure) is particularly suitable for generalization for a number of both practical and conceptual considerations.
-Our preliminary work shows that there are three, possible four, distinct set of requirements for an ensemble space $\Ens$, the set of ensemble given by a physical theory.
+The goal is to find a very limited set of axioms to define the entropic structure of a physical theory, which can be justified from physical considerations, show that it imposes a geometric structure on a generic space of statistical ensembles and that it recovers the standard geometric structures of classical and quantum mechanics.
+Furthermore, the same structure should be connected to topological and measure theoretic structures used in physics.
+
+%and Other that are required by any physical theory to define its entropic structure, which  and construct a general connection to geometry, calculus and topology, thus giving a generalized precise mathematical characterization of physical theories.
+In our numerous conversations with mathematicians, physicists and philosophers, it is clear that such work is sorely needed, and, since it lies at the intersection of different fields, it requires an interdisciplinary approach.
+
+We want to stress that finding the correct starting point is not just a mathematical issue, but also a physics problem.
+Let us explore this problem through an example.
+The space $M_1([0,1])$ of probability measures defined over the $[0,1]$ interval is an infinite dimensional vector space.
+As Aliprantis and Border note: ``Since there is more than one topology of interest on an infinite dimensional space, the choice of topology is a key modeling decision that can have economic as well as technical consequences.'' Unbeknowst to many physicists, this is true for physics as well.
+If we choose a weak topology, we are saying the convergence of expectation values of all continuous functions is enough to determine convergence of ensembles.
+While this may seem plausible at first glance, we can find a weakly convergence sequence for which the entropy does not converge.
+If the entropic structure of a physical theory is really equivalent to its geometric structure, not only the entropy should be continuous, but it is likely that the topology should be the one generated by the entropic structure.
+Ultimately, answering these questions means understanding both the physical and mathematical consequences of those decisions.
+
+Our preliminary work convinced us that the notion of statistical ensemble (i.e.
+the collection of all outputs of a preparation procedure) seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
+In terms of the math, the structure of statistical ensembles of both classical and quantum mechanics can be made to look very similar.
+From the experimental side, statistical ensembles are what we produce and measure experimentally.
+On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form ``whenever I prepare a system according to this procedure, I will find this outcome.'' That is, physical laws are relationships between ensembles.
+Therefore, it makes sense to require that a physical theory must \emph{at least} posit a space of ensembles.
+Our preliminary work shows that there are three, possibly four, distinct set of requirements for an ensemble space $\Ens$, the set of ensemble given by a physical theory.
 
@@
 On physical grounds, the mixing operation must be continuous in the natural topology defined by experimental verifiability, much like addition and scalar multiplication are continuous in a topological vector space, leading to the idea of a topological convex space.
-Unfortunately, we haven't found such a structure already studied in the literature.
-While it would be extremely useful for us to study these spaces on their own merit, we choose to concentrate on problems where our expertise in the mathematical foundations of physical theories is more critical, as in the next requirement.
+While we haven't found such a structure already studied in the literature, and it would extremely useful to us to see what can be generalized from topological vector spaces to topological convex spaces, this is a fairly well defined mathematical question at this point, and we believe that our expertise in the mathematical foundations of physical theories is better spent in continuing finding other well defined starting points.
 
-The third requirement is characterizing the variability of elements within a statistical ensemble.
-Though the instance within ensembles are never formalized, the variability within an ensemble is physically significant as it will influence, for example, the variability of statistical measurements.
-As we showed in previous work, entropy quantifies the variability of the elements within a distribution, and which gives a physically meaningful starting point.
-The proposed work is to find the exact specification of the axiom to see whether it can also guarantee that ensemble spaces must be metric spaces.
+The third requirement, central to the work, is characterizing the variability of elements within a statistical ensemble.
+Though we imagine ensembles to be a collection of instances, these instances are never formalized as they need to be left unspecified to leave the theory general.
+However, the variability of the elements within an ensemble is physically significant as it will influence, for example, the variability of statistical measurements.
+As we showed in previous work, entropy quantifies the variability of the elements within a distribution, and which gives us a physically meaningful starting point.
+The primary goal of the proposed work, then, is to find what axiomatic properties must this variability indicator have, and to understand what are their conceptual, physical and mathematical implications.
 
-The hardest conceptual problem is that entropy is defined differently in each space.
+The hardest conceptual problem we found is that the entropy is defined differently in each space.
 In a classical discrete space, the entropy is $-\sum p_i \log p_i$; in classical statistical mechanics is $- \int \rho \log \rho d\mu$ where $\mu$ is the Liouville measure; in quantum mechanics is $-\tr(\rho \log \rho)$.
@@
 Our approach is to simply characterize the entropy by its properties with respect to the topology and the convex structure.
-In our preliminary properties, for example, the entropy $S : \Ens \to \mathbb{R}$ has the following properties:
+In our preliminary work, for example, the entropy $S : \Ens \to \mathbb{R}$ has the following properties:
 \begin{itemize}
@@
 The proof is general, meaning that if we are right that those axioms must hold for any physical theory, then the Shannon entropy is of truly general applicability.
-Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is valid with an equal.
+Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is valid with an equal, and can be used to calculate the entropy from that decomposition.
 That allows to recover the expression for the entropy in the three different cases.
@@
 Our preliminary work shows that an entropy satisfying those axioms forces the convex space to be cancellative, meaning that it embeds into a vector space.
-That is, if we pick an internal point $\ens \in \Ens$, we can define a vector space $V_{\ens}$ such that $\Ens \subset V_{\ens}$.
-If we pick to different internal points $\ens \in \Ens$, we can show that there is an invertible affine map between $V_{\ens[a]}$ is $V_{\ens[b]}$.
+That is, if we pick an internal point $\ens \in \Ens$, we can define a vector space $V_{\ens}$ such that there exists an embedding $\iota : \Ens \hookrightarrow  V_{\ens}$ such that $\iota(p \ens[a] + \bar{p} \ens[b]) = p \iota(\ens[a]) + \bar{p} \iota(\ens[b])$ and $\iota(\ens) = 0$.
+If we pick two different internal points $\ens[a],\ens[b] \in \Ens$, we can show that there is an invertible affine map between $V_{\ens[a]}$ is $V_{\ens[b]}$.
 That is, the ensemble space embeds in an affine space.
@@
 
-We can also show that the entropy can be used to define geometric structures directly in the ensemble space.
+We can also show that the entropy can be used to define geometric structures directly on the ensemble space.
 For example, we can define the mixing entropy as $MS(\ens[a], \ens[b]) = S\left(\frac{1}{2}\ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2} S(\ens[a]) + \frac{1}{2} S(\ens[b])\right)$.
@@
 The JSD is the square of a distance function in both classical and quantum information theory, though our current axioms do not guarantee $MS$ to be so.
+It is an open question whether enough physical constraints exist for that to be the case.
 The $MS$ can be used to create open balls, and show that the topology must be Hausdorff.
+An interesting conjecture is whether an Hausdorff second countable topological convex space is metrizable, given that this is true for topological vector spaces.
 The current axioms do not guarantee the open balls to be convex.
+If we were to find sufficient physically motivated conditions that imply that the entropic open balls are convex and generate the topology, as we suspect it should be, the topology would be locally convex and therefore the entropic structure also gives us enough structure to define calculus.
 
@@
 This will be a differentiable manifold.
-On a chart whose coordinates are linear with respect to the affine structure we can define $g_{\ens}( v, w ) = - \frac{\partial^2 S}{\partial \ens ^2}( v, w )$ where $v$ and $w$ are elements of the tangent space at $\ens \in \Ens \subset V_{\ens[a]}$.
-Since all Hessian are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor.
+On a linear coordinate chart we can define $g_{\ens}( v, w ) = - \frac{\partial^2 S}{\partial \ens ^2}( v, w )$ where $v$ and $w$ are elements of the tangent space at $\ens \in \Ens \subset V_{\ens[a]}$.
+Since all Hessian are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is an Hessian manifold.
 In classical and quantum mechanics it recovers the Fisher information metric.
-Note that, if were to show continuous embedding into a locally convex topological vector space, this construction can be extended to the infinite dimensional case.
+The ultimate goal would be to show that, in general case, the embedding is continuous into a locally convex topological vector space so that this construction can be extended to the infinite dimensional case.
+This would essentially show that the entropy provides, in general, the geometric structure of the state space of any physical theory.
 
-understand whether the existence of a physically meaningful entropy function is enough to guarantee that the space of statistical ensembles of any physical theory is a convex subset of a locally convex topological vector space equipped with a metric.
-That is, we should not simply ``declare'' it to be so, but make sure that the mathematical starting points are justified by physical arguments, so that we can
+The fourth structure, which is left for future work, is the generalization of the Lie brackets given by the Poisson brackets in the classical case and the commutators in the quantum case.
+These describe the relationship between the generators of transformations that leave the geometric structure (i.e.
+symplectic form and inner product respectively) unchanged, which means, as outlined before, is equivalent to the geometric structure.
+That is, it describes the group of deterministic and reversible motions, those that leave the entropy unchanged.
+Given the tight connection between the entropic structure and the Lie algebraic structure, it is unclear whether this will be recovered automatically once the entropic structure is correctly axiomatized, or whether it will require additional axioms.
+For this reason, we postpone the study of this aspect.
+
+Having given an overview of the specifics of the problems, the proposed work consists of first finalizing and publishing some of the preliminary results, which include the equivalence of geometric and entropic structures in physical theories, the ability to recover the Shannon entropy in the generalized case and the three different entropies in the discrete/continuous/quantum cases.
+At the same time, explore whether the axiomatic can be expanded, without sacrificing potentially interesting physical cases, to recover potentially useful properties, such as the metrizability and local convexity of the topology generated by the entropy.
 

### 2025-10-04 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
2025NSF: typos and minor corrections

@@
 As part of our larger project, Assumptions of Physics, we aim to develop an axiomatic basis for states and processes that can be equally applied to different theories, such as classical and quantum mechanics.
-Different theory are then recovered as specializations of the more general structure.
+Different theories are then recovered as specializations of the more general structure.
 
@@
 In classical mechanics, a single degree of freedom is a two dimensional symplectic manifold.
-In this case, symplectic form $\omega$ is a volume form and therefore specifying $\omega$ is equivalent to specifying the Liouville measure $\mu(U) = \int_U \omega$.
+In this case, the symplectic form $\omega$ is a volume form and therefore specifying $\omega$ is equivalent to specifying the Liouville measure $\mu(U) = \int_U \omega$.
 In classical statistical mechanics, the entropy of a statistical ensemble is given by $S(\rho) = - \int \rho \log \rho d\mu$ where $\rho$ is the probability density associated to the ensemble.
-For a uniform distribution with support $U$, the entropy reduce to $S(\rho_U) = \log \mu(U)$.
+For a uniform distribution with support $U$, the entropy reduces to $S(\rho_U) = \log \mu(U)$.
 Since the logarithm is an invertible function, one can recover the measure from the entropy.
@@
 \end{equation}
-Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, mathematically $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi,\psi)}}{2},\frac{1 - \sqrt{p(\phi,\psi)}}{2} \right)$ where $I(x,y) = -x \log x -y \log y$ is the Shannon entropy.
-Therefore the entropy is an invertible function of $p(\psi,\phi)$.
+Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, mathematically $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has von Neumann entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi|\psi)}}{2},\frac{1 - \sqrt{p(\phi|\psi)}}{2} \right)$ where $I(x,y) = -x \log x -y \log y$ is the Shannon entropy.
+Therefore the entropy is an invertible function of $p(\phi|\psi)$.
 This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the state space.
 
-While the states spaces of classical (i.e.
+While the state spaces of classical (i.e.
 symplectic manifolds) and quantum mechanics (i.e.
 projective Hilbert spaces) are rather different, the space of ensembles of both have a similar structure.
-They are both convex subsets of a topological vector space, equipped with a strictly convex entropy function.
+They are both convex subsets of a topological vector space, equipped with a strictly concave entropy function.
 The core idea, then, is to use ensembles as a way to generalize the notion of physical states, and the entropy to characterize the geometry in a way that is independent of the particulars of each physical theory.
 In the past year we have done significant preliminary work that shows that not only this approach is possible, but is extremely productive.
-The proposed work is to turn those preliminary insight in an axiomatic specification that is fully physically motivated.
+The proposed work is to turn those preliminary insights in an axiomatic specification that is fully physically motivated.
 If successful, such an approach will provide the following benefits:
@@
 	\item provide a generalized structure for information theory, possibly allowing future generalization of results from both classical and quantum information theory
-	\item provide useful constrains to more precisely formulate physical theories in the infinite dimensional case.
+	\item provide useful constraints to more precisely formulate physical theories in the infinite dimensional case.
 \end{enumerate}
@@
 The space $M_1([0,1])$ of probability measures defined over the $[0,1]$ interval is an infinite dimensional vector space.
-As Aliprantis and Border note: ``Since there is more than one topology of interest on an infinite dimensional space, the choice of topology is a key modeling decision that can have economic as well as technical consequences.'' Unbeknowst to many physicists, this is true for physics as well.
+As Aliprantis and Border note: (add reference) ``Since there is more than one topology of interest on an infinite dimensional space, the choice of topology is a key modeling decision that can have economic as well as technical consequences.'' Unbeknowst to many physicists, this is true for physics as well.
 If we choose a weak topology, we are saying the convergence of expectation values of all continuous functions is enough to determine convergence of ensembles.
-While this may seem plausible at first glance, we can find a weakly convergence sequence for which the entropy does not converge.
+While this may seem plausible at first glance, we can find a weakly convergent sequence for which the entropy does not converge.
 If the entropic structure of a physical theory is really equivalent to its geometric structure, not only the entropy should be continuous, but it is likely that the topology should be the one generated by the entropic structure.
@@
 From the experimental side, statistical ensembles are what we produce and measure experimentally.
-On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form ``whenever I prepare a system according to this procedure, I will find this outcome.'' That is, physical laws are relationships between ensembles.
+On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form ``whenever I prepare a system according to this procedure, I will find this range of outcomes.'' That is, physical laws are relationships between ensembles.
 Therefore, it makes sense to require that a physical theory must \emph{at least} posit a space of ensembles.
@@
 statement associated with an experimental test without termination guarantees) and Borel sets.
-Real and integer numbers, for example, are recovered from an idealized model representing references of a reference system.
+(add citation) Real and integer numbers, for example, are recovered from an idealized model representing references of a reference system.
 The key insight is that the set theoretic operations represent a two-valued logic system (i.e.
 truth or falsehood of statements) while the topology keeps track of a three-valued logic system (i.e.
-interior, exterior and boundary represent respectively successful termination, unsuccessful termination and non-termination not terminate).
+interior, exterior and boundary represent respectively successful termination, unsuccessful termination and non-termination).
 This work gives us a solid foundation on which to build, and guarantees that an ensemble space $\Ens$ is, at least, a $\mathsf{T}_0$ second countable topological space.
@@
 This is insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs).
-The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to physically untenable cases).
+The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases).
 Rather, we let the topology do the work.
@@
 On physical grounds, the mixing operation must be continuous in the natural topology defined by experimental verifiability, much like addition and scalar multiplication are continuous in a topological vector space, leading to the idea of a topological convex space.
-While we haven't found such a structure already studied in the literature, and it would extremely useful to us to see what can be generalized from topological vector spaces to topological convex spaces, this is a fairly well defined mathematical question at this point, and we believe that our expertise in the mathematical foundations of physical theories is better spent in continuing finding other well defined starting points.
+While we haven't found such a structure already studied in the literature, and it would be extremely useful to us to see what can be generalized from topological vector spaces to topological convex spaces, this is a fairly well defined mathematical question at this point, and we believe that our expertise in the mathematical foundations of physical theories is better spent in continuing finding other well defined starting points.
 
@@
 How can we give a single definition for these very different expressions?
-Note that the expressions are different precisely because they are in terms of the specific representation of ensembles in specific theory.
+Note that the expressions are different precisely because they are in terms of the specific representation of ensembles in the specific theory.
 Our approach is to simply characterize the entropy by its properties with respect to the topology and the convex structure.
-In our preliminary work, for example, the entropy $S : \Ens \to \mathbb{R}$ has the following properties:
+In our preliminary work, for example, the entropy $S : \Ens \to \mathbb{R}$ has currently the following properties:
 \begin{itemize}
@@
 These properties are justified by the physics.
-For example, if we are performing a statistical mixture between two ensemble, the average variability cannot decrease, which justifies the concavity.
-These relatively simple properties allow us, for example, to show that $I(p,\bar{p})$, the maximum entropy increase during mixing, must be $ - \kappa \left(p \log p + \bar{p} \log \bar{p}\right)$, the Shannon entropy.
+For example, if we are performing a statistical mixture between two ensembles, the average variability cannot decrease, which justifies the concavity.
+These relatively simple properties allow us, for example, to show that $I(p,\bar{p})$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy $ - \left(p \log p + \bar{p} \log \bar{p}\right)$.
 The proof is general, meaning that if we are right that those axioms must hold for any physical theory, then the Shannon entropy is of truly general applicability.
@@
 That is, if we pick an internal point $\ens \in \Ens$, we can define a vector space $V_{\ens}$ such that there exists an embedding $\iota : \Ens \hookrightarrow  V_{\ens}$ such that $\iota(p \ens[a] + \bar{p} \ens[b]) = p \iota(\ens[a]) + \bar{p} \iota(\ens[b])$ and $\iota(\ens) = 0$.
-If we pick two different internal points $\ens[a],\ens[b] \in \Ens$, we can show that there is an invertible affine map between $V_{\ens[a]}$ is $V_{\ens[b]}$.
-That is, the ensemble space embeds in an affine space.
-We can also show that it is ``directionally bounded,'' meaning its intersection with any affine line is finite.
-It is an open question whether the axioms are enough to prove that the embedding is continuous in a topological vector space, and whether the set is bounded.
+If we pick two different internal points $\ens[a],\ens[b] \in \Ens$, we can show that there is an invertible affine map between $V_{\ens[a]}$ and $V_{\ens[b]}$.
+That is, the ensemble space should be more properly thought as embedding in a affine space.
+We can also show that it is ``directionally bounded,'' meaning its intersection with any affine line is finite interval.
+It is an open question whether the current axioms are enough to prove that the embedding is continuous in a topological vector space, and whether the ensemble space is a bounded subset.
 
@@
 It is almost a distance function, as it does not satisfy the triangle inequality, but it recovers the Jensen-Shannon divergence (JSD) in both classical and quantum mechanics.
-The JSD is the square of a distance function in both classical and quantum information theory, though our current axioms do not guarantee $MS$ to be so.
+While JSD is the square of a distance function in both classical and quantum information theory, our current axioms do not guarantee $MS$ to have this property in general.
 It is an open question whether enough physical constraints exist for that to be the case.
 The $MS$ can be used to create open balls, and show that the topology must be Hausdorff.
-An interesting conjecture is whether an Hausdorff second countable topological convex space is metrizable, given that this is true for topological vector spaces.
+An interesting conjecture is whether a Hausdorff second countable topological convex space is metrizable, given that this is true for topological vector spaces.
 The current axioms do not guarantee the open balls to be convex.
-If we were to find sufficient physically motivated conditions that imply that the entropic open balls are convex and generate the topology, as we suspect it should be, the topology would be locally convex and therefore the entropic structure also gives us enough structure to define calculus.
+If we were to find sufficient physically motivated conditions that imply that the entropic open balls are convex and generate the topology, as we suspect it should be, the topology would be locally convex and therefore the entropic structure also may give us enough structure to define calculus.
 

### 2025-10-07 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Edits before sending draft to Alejandro

@@
 
-A core insight that is guiding this enterprise is that geometric structures in both classical and quantum mechanics are equivalent to entropic structures.
-In classical mechanics, a single degree of freedom is a two dimensional symplectic manifold.
+A core insight guiding this enterprise is that geometric structures in both classical and quantum mechanics are equivalent to entropic structures.
+In classical mechanics, a single degree of freedom is a two-dimensional symplectic manifold.
 In this case, the symplectic form $\omega$ is a volume form and therefore specifying $\omega$ is equivalent to specifying the Liouville measure $\mu(U) = \int_U \omega$.
@@
 That is, specifying the entropy associated to all uniform distributions allows us to reconstruct the Liouville measure, and therefore the symplectic form.
-In the general case, Darboux coordinates allow, at least locally, to factorize the space into multiple independent degrees of freedom, where the volume is the product of the corresponding areas.
+In the general case, Darboux coordinates allow us, at least locally, to factorize the space into multiple independent degrees of freedom, where the volume is the product of the corresponding areas.
 The corresponding entropic property is that the entropy for the joint distribution is the sum of the marginals if the independent degrees of freedom are independently distributed.
-Therefore, if the distribution of all possible marginals are specified, the symplectic form can be recovered.
+Therefore, if the distributions of all possible marginals are specified, the symplectic form can be recovered.
 
@@
 
-While the state spaces of classical (i.e.
-symplectic manifolds) and quantum mechanics (i.e.
-projective Hilbert spaces) are rather different, the space of ensembles of both have a similar structure.
+While the state spaces of classical (i.e.~symplectic manifolds) and quantum mechanics (i.e.~projective Hilbert spaces) are rather different, the space of ensembles of both have a similar structure.
 They are both convex subsets of a topological vector space, equipped with a strictly concave entropy function.
 The core idea, then, is to use ensembles as a way to generalize the notion of physical states, and the entropy to characterize the geometry in a way that is independent of the particulars of each physical theory.
-In the past year we have done significant preliminary work that shows that not only this approach is possible, but is extremely productive.
+In the past year, we have done significant preliminary work that shows that not only is this approach possible, but it is extremely productive.
 The proposed work is to turn those preliminary insights in an axiomatic specification that is fully physically motivated.
@@
 	\item provide a generalized structure for information theory, possibly allowing future generalization of results from both classical and quantum information theory
-	\item provide useful constraints to more precisely formulate physical theories in the infinite dimensional case.
+	\item provide useful constraints to more precisely formulate physical theories in the infinite-dimensional case.
 \end{enumerate}
@@
 \section{Proposed research program}
+\subsection{Intellectual merit}
 
-The goal is to find a very limited set of axioms to define the entropic structure of a physical theory, which can be justified from physical considerations, show that it imposes a geometric structure on a generic space of statistical ensembles and that it recovers the standard geometric structures of classical and quantum mechanics.
+The goal is to find a very limited set of axioms to define the entropic structure of a physical theory that can be justified from physical considerations, show that this set of axioms imposes a geometric structure on a generic space of statistical ensembles and that it recovers the standard geometric structures of classical and quantum mechanics.
 Furthermore, the same structure should be connected to topological and measure theoretic structures used in physics.
@@
 Let us explore this problem through an example.
-The space $M_1([0,1])$ of probability measures defined over the $[0,1]$ interval is an infinite dimensional vector space.
-As Aliprantis and Border note: (add reference) ``Since there is more than one topology of interest on an infinite dimensional space, the choice of topology is a key modeling decision that can have economic as well as technical consequences.'' Unbeknowst to many physicists, this is true for physics as well.
+The space $M_1([0,1])$ of probability measures defined over the $[0,1]$ interval is an infinite-dimensional vector space.
+As Aliprantis and Border note in the context of their own field of economics [add reference],  ``Since there is more than one topology of interest on an infinite dimensional space, the choice of topology is a key modeling decision that can have economic as well as technical consequences.'' Unbeknowst to many physicists, this is true for physics as well.
 If we choose a weak topology, we are saying the convergence of expectation values of all continuous functions is enough to determine convergence of ensembles.
 While this may seem plausible at first glance, we can find a weakly convergent sequence for which the entropy does not converge.
-If the entropic structure of a physical theory is really equivalent to its geometric structure, not only the entropy should be continuous, but it is likely that the topology should be the one generated by the entropic structure.
+If the entropic structure of a physical theory is really equivalent to its geometric structure, not only should the entropy be continuous, but it is likely that the topology should be the one generated by the entropic structure.
 Ultimately, answering these questions means understanding both the physical and mathematical consequences of those decisions.
 
-Our preliminary work convinced us that the notion of statistical ensemble (i.e.
-the collection of all outputs of a preparation procedure) seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
+Our preliminary work convinced us that the notion of statistical ensemble (i.e.~the collection of all outputs of a preparation procedure) seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
 In terms of the math, the structure of statistical ensembles of both classical and quantum mechanics can be made to look very similar.
@@
 On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form ``whenever I prepare a system according to this procedure, I will find this range of outcomes.'' That is, physical laws are relationships between ensembles.
-Therefore, it makes sense to require that a physical theory must \emph{at least} posit a space of ensembles.
-Our preliminary work shows that there are three, possibly four, distinct set of requirements for an ensemble space $\Ens$, the set of ensemble given by a physical theory.
+Therefore, it makes sense to require that any physical theory must \emph{at least} posit a space of ensembles.
+Our preliminary work shows that there are three, possibly four, distinct sets of requirements for an ensemble space $\Ens$, the set of ensembles given by a physical theory.
+The work proposed here focuses on further development of the third.
 
 The first requirement is that of experimental verifiability of the theory.
-This has been thoroughly investigated in our previous work that established a precise correspondence between experimental verifiable statements (i.e.
-statement associated with an experimental test that terminates successfully in finite time if and only if the statement is true) and the open set of a ``natural'' topology, and between theoretical statements (i.e.
-statement associated with an experimental test without termination guarantees) and Borel sets.
-(add citation) Real and integer numbers, for example, are recovered from an idealized model representing references of a reference system.
-The key insight is that the set theoretic operations represent a two-valued logic system (i.e.
-truth or falsehood of statements) while the topology keeps track of a three-valued logic system (i.e.
-interior, exterior and boundary represent respectively successful termination, unsuccessful termination and non-termination).
-This work gives us a solid foundation on which to build, and guarantees that an ensemble space $\Ens$ is, at least, a $\mathsf{T}_0$ second countable topological space.
+This has been thoroughly investigated in our previous work that established a precise correspondence between experimental verifiable statements (i.e.~statements associated with an experimental test that terminates successfully in finite time if and only if the statement is true) and the open set of a ``natural'' topology, and between theoretical statements (i.e.~statement associated with an experimental test without termination guarantees) and Borel sets~\cite{aop-book,aop-topExpDisting}.
+Real and integer numbers, for example, are recovered from an idealized model representing references of a reference system.
+The key insight is that the set-theoretic operations represent a two-valued logic system (i.e.~truth or falsehood of statements) while the topology keeps track of a three-valued logic system (i.e.~interior, exterior and boundary represent respectively successful termination, unsuccessful termination and non-termination).
+This work gives us a solid foundation on which to build, and guarantees that an ensemble space $\Ens$ is, at least, a $\mathsf{T}_0$ second-countable topological space.
 
@@
 This gives the ensemble space a convex structure as $p \ens[a] + \bar{p} \ens[b] \in \Ens$ for all $\ens[a], \ens[b] \in \Ens$ and $p \in [0,1]$.
-This is insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs).
+This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs) [add refs].
 The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases).
@@
 On physical grounds, the mixing operation must be continuous in the natural topology defined by experimental verifiability, much like addition and scalar multiplication are continuous in a topological vector space, leading to the idea of a topological convex space.
-While we haven't found such a structure already studied in the literature, and it would be extremely useful to us to see what can be generalized from topological vector spaces to topological convex spaces, this is a fairly well defined mathematical question at this point, and we believe that our expertise in the mathematical foundations of physical theories is better spent in continuing finding other well defined starting points.
+We haven't found such a structure already studied in the literature.
+While it would be extremely useful to us to see what can be generalized from topological vector spaces to topological convex spaces, we do not believe that domain knowledge in physics is specifically valuable here and leave this work to others, as we believe that our expertise in the mathematical foundations of physical theories is better spent in continuing to find other well defined starting points.
 
-The third requirement, central to the work, is characterizing the variability of elements within a statistical ensemble.
+The third requirement, central to the proposed work, is characterizing the variability of elements within a statistical ensemble.
 Though we imagine ensembles to be a collection of instances, these instances are never formalized as they need to be left unspecified to leave the theory general.
 However, the variability of the elements within an ensemble is physically significant as it will influence, for example, the variability of statistical measurements.
-As we showed in previous work, entropy quantifies the variability of the elements within a distribution, and which gives us a physically meaningful starting point.
-The primary goal of the proposed work, then, is to find what axiomatic properties must this variability indicator have, and to understand what are their conceptual, physical and mathematical implications.
+As we showed in previous work~\cite{Carcassi:2021}, entropy quantifies the variability of the elements within a distribution, which gives us a physically meaningful starting point.
+The primary goal of the proposed work, then, is to find what axiomatic properties this variability indicator must have, and to understand their conceptual, physical and mathematical implications.
 
 The hardest conceptual problem we found is that the entropy is defined differently in each space.
-In a classical discrete space, the entropy is $-\sum p_i \log p_i$; in classical statistical mechanics is $- \int \rho \log \rho d\mu$ where $\mu$ is the Liouville measure; in quantum mechanics is $-\tr(\rho \log \rho)$.
+In a classical discrete space, the entropy is $-\sum p_i \log p_i$; in classical statistical mechanics it is $- \int \rho \log \rho d\mu$ where $\mu$ is the Liouville measure; in quantum mechanics it is $-\tr(\rho \log \rho)$.
 How can we give a single definition for these very different expressions?
@@
 Our approach is to simply characterize the entropy by its properties with respect to the topology and the convex structure.
-In our preliminary work, for example, the entropy $S : \Ens \to \mathbb{R}$ has currently the following properties:
+In our preliminary work, for example, the axiom defining entropy $S : \Ens \to \mathbb{R}$ currently has the following properties:
 \begin{itemize}
@@
 	\item Strict concavity: $S(p\ens[a] + \bar{p} \ens[b]) \geq p S(\ens[a]) + \bar{p} S(\ens[b])$ with the equality holding if and only if $\ens[a] = \ens[b]$
-	\item Upper variability bound: there exists a universal function $I(p_1, p_2)$ (i.e.
-the same for all ensemble spaces) such that $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p, \bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])$; if the equality holds, $\ens[a]$ and $\ens[b]$ are orthogonal, noted $\ens[a] \ortho \ens[b]$
+	\item Upper variability bound: there exists a universal function $I(p_1, p_2)$ (i.e.~the same for all ensemble spaces) such that $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p, \bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])$; if the equality holds, $\ens[a]$ and $\ens[b]$ are orthogonal, noted $\ens[a] \ortho \ens[b]$
 	\item Mixtures preserve orthogonality: $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$
@@
 For example, if we are performing a statistical mixture between two ensembles, the average variability cannot decrease, which justifies the concavity.
-These relatively simple properties allow us, for example, to show that $I(p,\bar{p})$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy $ - \left(p \log p + \bar{p} \log \bar{p}\right)$.
+These relatively simple properties allow us, for example, to show that $I(p,\bar{p})$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy~\cite{Shannon} $ - \left(p \log p + \bar{p} \log \bar{p}\right)$.
 The proof is general, meaning that if we are right that those axioms must hold for any physical theory, then the Shannon entropy is of truly general applicability.
-Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is valid with an equal, and can be used to calculate the entropy from that decomposition.
-That allows to recover the expression for the entropy in the three different cases.
+Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is an equality and can be used to calculate the entropy from that decomposition.
+This allows us to recover the expressions for the entropy in the three different cases, classical discete, classical continuous, and quantum.
 
-Our preliminary work shows that an entropy satisfying those axioms forces the convex space to be cancellative, meaning that it embeds into a vector space.
+Our preliminary work shows that an entropy satisfying those properties forces the convex space to be cancellative, meaning that it embeds into a vector space.
 That is, if we pick an internal point $\ens \in \Ens$, we can define a vector space $V_{\ens}$ such that there exists an embedding $\iota : \Ens \hookrightarrow  V_{\ens}$ such that $\iota(p \ens[a] + \bar{p} \ens[b]) = p \iota(\ens[a]) + \bar{p} \iota(\ens[b])$ and $\iota(\ens) = 0$.
 If we pick two different internal points $\ens[a],\ens[b] \in \Ens$, we can show that there is an invertible affine map between $V_{\ens[a]}$ and $V_{\ens[b]}$.
-That is, the ensemble space should be more properly thought as embedding in a affine space.
-We can also show that it is ``directionally bounded,'' meaning its intersection with any affine line is finite interval.
-It is an open question whether the current axioms are enough to prove that the embedding is continuous in a topological vector space, and whether the ensemble space is a bounded subset.
+That is, the ensemble space should be more properly thought as embedding into an affine space.
+We can also show that it is ``directionally bounded,'' meaning its intersection with any affine line is a finite interval.
+It is an open question whether the currently defined properties are enough to prove that the embedding is continuous in a topological vector space, and whether the ensemble space is a bounded subset.
 
@@
 If we express the entropy in bits, the mixing entropy satisfies the following:
-\begin{enumerate}
+\begin{itemize}
 	\item non-negativity: $MS(\ens[a], \ens[b]) \geq 0$
@@
 	\item symmetry: $MS(\ens[a], \ens[b]) = MS(\ens[b], \ens[a])$
-\end{enumerate}
-It is almost a distance function, as it does not satisfy the triangle inequality, but it recovers the Jensen-Shannon divergence (JSD) in both classical and quantum mechanics.
-While JSD is the square of a distance function in both classical and quantum information theory, our current axioms do not guarantee $MS$ to have this property in general.
+\end{itemize}
+It is almost a distance function, as it does not satisfy the triangle inequality, but it recovers the Jensen-Shannon divergence (JSD) in both classical [add ref] and quantum [add ref] mechanics.
+While the JSD is the square of a distance function in both classical and quantum information theory, our current axioms do not guarantee $MS$ to have this property in general.
 It is an open question whether enough physical constraints exist for that to be the case.
-The $MS$ can be used to create open balls, and show that the topology must be Hausdorff.
-An interesting conjecture is whether a Hausdorff second countable topological convex space is metrizable, given that this is true for topological vector spaces.
+The $MS$ can be used to create open balls and show that the topology must be Hausdorff.
+An interesting conjecture is whether a Hausdorff second-countable topological convex space is metrizable, given that this is true for topological vector spaces.
 The current axioms do not guarantee the open balls to be convex.
-If we were to find sufficient physically motivated conditions that imply that the entropic open balls are convex and generate the topology, as we suspect it should be, the topology would be locally convex and therefore the entropic structure also may give us enough structure to define calculus.
+If we were to find sufficient physically motivated conditions that imply that the entropic open balls are convex and generate the topology, as we suspect should be the case, the topology would be locally convex and therefore the entropic structure also may give us enough structure to define calculus.
 
-If we assume we are in finite dimensions and that the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite which allows us to define a metric tensor.
+If we assume we are in finite dimensions and that the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
 Let $V_{\ens[a]}$ be the embedding vector space with $\ens[a] \in \Ens$ as origin.
@@
 On a linear coordinate chart we can define $g_{\ens}( v, w ) = - \frac{\partial^2 S}{\partial \ens ^2}( v, w )$ where $v$ and $w$ are elements of the tangent space at $\ens \in \Ens \subset V_{\ens[a]}$.
-Since all Hessian are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is an Hessian manifold.
+Since all Hessians are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is a Hessian manifold.
 In classical and quantum mechanics it recovers the Fisher information metric.
-The ultimate goal would be to show that, in general case, the embedding is continuous into a locally convex topological vector space so that this construction can be extended to the infinite dimensional case.
+The ultimate goal would be to show that, in the general case, the embedding is continuous into a locally convex topological vector space so that this construction can be extended to the infinite-dimensional case.
 This would essentially show that the entropy provides, in general, the geometric structure of the state space of any physical theory.
 
-The fourth structure, which is left for future work, is the generalization of the Lie brackets given by the Poisson brackets in the classical case and the commutators in the quantum case.
-These describe the relationship between the generators of transformations that leave the geometric structure (i.e.
-symplectic form and inner product respectively) unchanged, which means, as outlined before, is equivalent to the geometric structure.
-That is, it describes the group of deterministic and reversible motions, those that leave the entropy unchanged.
-Given the tight connection between the entropic structure and the Lie algebraic structure, it is unclear whether this will be recovered automatically once the entropic structure is correctly axiomatized, or whether it will require additional axioms.
+The fourth requirement, which is left for future work, is the generalization of the Lie brackets for a generic ensemble space, given by the Poisson brackets in the classical case and the commutators in the quantum case.
+These describe the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged, which means, as outlined before, leaving the entropic structure unchanged.
+That is, it describes the group of deterministic and reversible motions, those that leave the entropy unchanged~\cite{aop-HamConsInfoEnt,aop-HamQuantInfo,aop-book}.
+Given the tight connection between the entropic structure and the Lie-algebraic structure, it is unclear whether this will be recovered automatically once the entropic structure is correctly axiomatized, or whether it will require additional axioms.
 For this reason, we postpone the study of this aspect.
@@
 Having given an overview of the specifics of the problems, the proposed work consists of first finalizing and publishing some of the preliminary results, which include the equivalence of geometric and entropic structures in physical theories, the ability to recover the Shannon entropy in the generalized case and the three different entropies in the discrete/continuous/quantum cases.
-At the same time, explore whether the axiomatic can be expanded, without sacrificing potentially interesting physical cases, to recover potentially useful properties, such as the metrizability and local convexity of the topology generated by the entropy.
+At the same time, we will explore whether the axiomatics can be expanded without sacrificing potentially interesting physical cases in order to recover potentially useful properties, such as the metrizability of the space and local convexity of the topology generated by the entropy.
 
 Notes:
-
-Generalized ensemble spaces + geometry defined from entropy; full set of axioms
 
@@
 
-Have: 1. Topological structure; 2. Convex structure; 3. Geometric structure.
-
-Open problems: 1. Entropic open balls can generate topology? 2. Are all axioms independent?   …
-
 To determine which fundamental structures/properties are needed for all physical theories.
@@
 
-Mention economists evidently aware of topological issues more than physicists.
-
-\section{Proposed research program}
-
-\subsection{Intellectual merit}
-
 \section{Summary of proposed research program}
+[Add brief recap]
 
@@
 Her separate line of theoretical research on the mathematical foundations of physics is relatively new, with her first publication in 2018~\cite{aop-phys-blueprint}.
-She has successfully split her efforts between her different research lines for the past six years, with her work related to the Assumptions of Physics project producing 14 peer-reviewed publications, 1 more submitted paper, and a published technical book since 2018.
+She has successfully split her efforts between her different research lines for the past seven years, with her work related to the Assumptions of Physics project producing 15 peer-reviewed publications, 1 more submitted paper, and a published technical book since 2018.
 She is moreover in the process of reducing the number of personnel she supervises on QCD research in order to have more time for her theoretical work---four of her Ph.D.~students in QCD completed their degrees in the last 18 months, and two more are expected to graduate before the start of the proposed project.
@@
 \\
-TODO: Add Alejandro assuming he agrees.
+TODO: Add Alejandro if he agrees.
 
@@
 
-\section{Broader impacts}
+\subsection{Broader impacts}
 In addition to disseminating the goals and results of the project through traditional means such as peer-reviewed publications and presentations at workshops and conferences, the goals and results of the project will be disseminated to wider audiences by leveraging existing social media and other outreach infrastructure that has already been developed for the broader Assumptions of Physics project, in particular, Carcassi's two YouTube channels, as described below in the section on qualifications of the research team.
@@
 Graduate student Tobias Thrien originally discovered the AoP project through Carcassi's YouTube channel while a Bachelor's student in Germany several years ago.
-He has since completed a Master's in theoretical cosmology in Germany and is currently a first-year Ph.D.~student at U-M, having come to Michigan specifically to work on the AoP project with the PI and Carcassi.
+He has since completed a Master's in theoretical cosmology in Germany and is currently a second-year Ph.D.~student at U-M, having come to Michigan specifically to work on the AoP project with the PI and Carcassi.
 Regarding the impact of posting open technical problems, our first collaborative paper that resulted from someone reaching out to us after seeing a video, on classical mechanics as the high-entropy limit of quantum mechanics, is currently under review at a journal~\cite{aop-classicallimit}.
 We were previously completely unaware of the researcher and his work at the University of Innsbruck, Austria.
+Carcassi's August 2025 video on ``The need for physical mathematics'' has already received 49k views and led directly to an invitation to speak at the American Mathematical Society conference in Washington, DC in January 2026.
 Carcassi additionally writes monthly blog-style essays related to the AoP project and advertised through social media platforms such as Facebook and LinkedIn.
@@
 
-TODO: Paragraph about impact of unifying disparate branches of physics, and eventual impact on physics education.\\
+Carcassi and Aidala will host an annual online school that includes an introduction to generalized ensemble spaces and the mathematical relationships between geometric and entropic structures, showcasing various results from the proposed work in a didactic manner.
+The school will consist of six hour-long lectures, two per day over three days, each followed by 30 minutes of questions and discussion, from 10 am - 1 pm U.S.
+Eastern Time, enabling individuals from across the Americas, Europe, Africa, and parts of Asia to participate at a reasonable hour.
+The school will be advertised primarily through social media.
+Registrants will be given a Zoom link for participation, and those who choose not to register will be able to follow a livestream on YouTube (on Carcassi's technical channel) and ask questions through the livestream chat.
+There will be no fee to participate.
+All lectures and discussion will also be recorded and available through YouTube later.
+With 27 registrants for the inaugural Assumptions of Physics online school in June 2024, this grew to 167 registrants in June 2025, mainly undergraduate and graduate students and recent graduates in physics and/or mathematics, but also a handful of professors and other senior researchers.
+Among the registrants were individuals from India, Bangladesh, Pakistan, Ethiopia, Egypt, Sudan, Bolivia, Brazil, and Singapore, in addition to North America and Europe.
+During the period of proposed work, we expect the school to reach several hundred individuals across similar demographics.
+
+Beyond the immediate research goals, the proposed work will serve to unify what are currently disparate branches of physics---classical and quantum, and aspects of statistical mechanics.
+Developing a common framework will pave the way for eventual new approaches to physics education, with less compartmentalized teaching across the core advanced undergraduate courses.
+Finding general mathematical structures that any physical theory must have will also help constrain the search for new physical theories.
+\\
 
 \noindent
-\textbf{Assessment of broader impacts:}  The reach and impact of videos related to the proposed work will be assessed by the number of views they receive and by other measures of audience engagement such as comments posted or subsequent email contacts.
+\textbf{Assessment of broader impacts:}  The reach and impact of videos related to the proposed work will be assessed by the number of views they receive and by other measures of audience engagement such as comments posted or subsequent email contacts, as well as potential talk invitations or proposals for collaboration.
+The reach and impact of the school will be assessed by the number of registrants and their demographics, as well as the subsequent views of the recorded lectures.
 The impact of unifying disparate branches of physics will need to be assessed on a longer-term timescale than that of the grant, but early indicators could include talk invitations and paper citations.
 \\
-TODO: Consider ways to assess engagement in more detail, per last year's reviews.
 
@@
 She was a 2019-20 U.S.~Fulbright Scholar and is a Fellow of the American Physical Society (APS).
-She served on the National Academy of Sciences Committee on a U.S.-Based Electron-Ion Collider Science Assessment 2016-18 and is currently serving on the U.S.~Nuclear Science Advisory Committee.
+She served on the National Academy of Sciences Committee on a U.S.-Based Electron-Ion Collider Science Assessment 2016-18 and on the U.S.~Nuclear Science Advisory Committee 2023-25.
 She served as an elected member of the Executive Committee of the APS Division of Nuclear Physics (DNP) 2021-23, and since 2021 she has served as a DNP Ally to ensure that DNP meetings are welcoming and inclusive environments.
-Much of her work has been in quantum chromodynamics; however, in 2015 she initiated the Assumptions of Physics project with her collaborator and husband Gabriele Carcassi, and since 2018 they have so far published together 14 peer-reviewed articles~\cite{aop-phys-blueprint,aop-topExpDisting,aop-HamQuantInfo,Carcassi:2021,aop-spacetimeStruct,aop-HamConsInfoEnt,Carcassi2021four,Carcassi:2022bpm,aop-HamPriv,aop-commonLogical,aop-action,aop-nogo,aop-nonaddmeas,aop-unphysHilbert} as well as a technical book on the mathematical foundations of physics~\cite{aop-book}, with one additional article currently under review and publicly available as a preprint~\cite{aop-classicallimit}.
+Much of her work has been in quantum chromodynamics; however, in 2015 she initiated the Assumptions of Physics project with her collaborator and husband Gabriele Carcassi, and since 2018 they have so far published together 15 peer-reviewed articles [one more to add]~\cite{aop-phys-blueprint,aop-topExpDisting,aop-HamQuantInfo,Carcassi:2021,aop-spacetimeStruct,aop-HamConsInfoEnt,Carcassi2021four,Carcassi:2022bpm,aop-HamPriv,aop-commonLogical,aop-action,aop-nogo,aop-nonaddmeas,aop-unphysHilbert} as well as a technical book on the mathematical foundations of physics~\cite{aop-book}, with one additional article currently under review and publicly available as a preprint~\cite{aop-classicallimit}.
 She is the PI on a grant from the John Templeton Foundation that supports the development of novel measure theoretic structures for quantum mechanics.
@@
 
-TODO: Add info for Alejandro.
+TODO: Add info for Alejandro if agrees.
 
@@
 He is the Co-PI on the Templeton Foundation grant with Aidala.
-Related to the AoP project, he has given seminars to audiences of physicists, mathematicians, and philosophers at U-M, Harvard, Caltech, the University of Pavia, Italy, the Munich Center for Mathematical Philosophy, Germany, and the Wolfram Institute, and he and Aidala gave a joint public talk for the Michigan State University Advanced Studies Gateway.
+Related to the AoP project, he has given seminars to audiences of physicists, mathematicians, and philosophers at places including Harvard, Caltech, Oxford (UK) the University of Pavia (Italy) the Munich Center for Mathematical Philosophy (Germany) and the Wolfram Institute.
+He and Aidala gave a joint public talk for the Michigan State University Advanced Studies Gateway, and he will give a U-M ``Saturday Morning Physics'' public talk in November 2025.
 His training in engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from theoretical computer science, information theory, systems theory, signal processing, and operations research.
@@
 \\
-TODO: Add Kansas, Saturday Morning Physics, invited talk at AMS meeting, also earlier Topology conference talk?
 
-Together, Aidala and Carcassi are uniquely positioned to advance the mathematical foundations of physics, with Aidala's traditional training in physics and traditional academic position complementing Carcassi's formal training in engineering and his university staff position, which for example gave him the time and flexibility to audit multiple graduate Mathematics courses at U-M.
+Together, Aidala and Carcassi [add Alejandro if agrees] are uniquely positioned to advance the mathematical foundations of physics, with Aidala's traditional training in physics and traditional academic position complementing Carcassi's formal training in engineering and his university staff position, which for example gave him the time and flexibility to audit multiple graduate Mathematics courses at U-M.
 For the AoP project, their published work so far has covered a wide range of topics.
@@
 Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds, and in 2025 the U-M Rackham Distinguished Graduate Mentor Award.
-She has served as the dissertation chair for 10 graduate students in Physics and Applied Physics who have already completed their doctorates, as co-chair for one in Applied and Indisciplinary Mathematics and another in Applied Physics, and is currently the dissertation chair for an additional 4 Ph.D.~students in Physics.
+She has served as the dissertation chair for 10 graduate students in Physics and Applied Physics who have already completed their doctorates, as co-chair with Uribe for one in Applied and Indisciplinary Mathematics and another in Applied Physics, and is currently the dissertation chair for an additional 4 Ph.D.~students in Physics.
 Aidala and Carcassi have also successfully published with former U-M Ph.D.~students in Mathematics~\cite{aop-spacetimeStruct} and Philosophy~\cite{aop-HamPriv} whose primary dissertation work was with other faculty members.
@@
 
-TODO: Add info on summer schools here?
-
 TODO: Add info for Alejandro.
 
-Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with more than 17,500 subscribers currently.
+Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with more than 17,600 subscribers currently.
 He posts a new video monthly, generally targeted at individuals with an advanced undergraduate background in physics and/or mathematics.
@@
 Given the breadth of the broader AoP project, there is no single community with the appropriate technical background across all relevant and potentially relevant areas of mathematics and physics.
-So far this ``crowd-sourcing'' approach has led to multiple instances of viewers with whom we have had no previous contact pointing us to valuable reference materials, and it recently led to our first coauthored paper with a researcher who reached out to us after seeing an open problem video, on ``Classical mechanics as the high-entropy limit of quantum mechanics''~\cite{aop-classicallimit}.
+So far this ``crowd-sourcing'' approach has led to multiple instances of viewers with whom we have had no previous contact pointing us to valuable reference materials, and in 2024 it led to our first coauthored paper with a researcher who reached out to us after seeing an open problem video, on ``Classical mechanics as the high-entropy limit of quantum mechanics''~\cite{aop-classicallimit}.
 We were previously completely unaware of the researcher and his work in Innsbruck, Austria.

### 2025-10-19 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
2025NSF: reworked the narrative

@@
 
+\iffalse
 Physics is currently a patchwork of different theories with no common underlying foundation.
@@
 Different theories are then recovered as specializations of the more general structure.
-
-A core insight guiding this enterprise is that geometric structures in both classical and quantum mechanics are equivalent to entropic structures.
-In classical mechanics, a single degree of freedom is a two-dimensional symplectic manifold.
-In this case, the symplectic form $\omega$ is a volume form and therefore specifying $\omega$ is equivalent to specifying the Liouville measure $\mu(U) = \int_U \omega$.
-In classical statistical mechanics, the entropy of a statistical ensemble is given by $S(\rho) = - \int \rho \log \rho d\mu$ where $\rho$ is the probability density associated to the ensemble.
-For a uniform distribution with support $U$, the entropy reduces to $S(\rho_U) = \log \mu(U)$.
-Since the logarithm is an invertible function, one can recover the measure from the entropy.
-That is, specifying the entropy associated to all uniform distributions allows us to reconstruct the Liouville measure, and therefore the symplectic form.
-In the general case, Darboux coordinates allow us, at least locally, to factorize the space into multiple independent degrees of freedom, where the volume is the product of the corresponding areas.
-The corresponding entropic property is that the entropy for the joint distribution is the sum of the marginals if the independent degrees of freedom are independently distributed.
-Therefore, if the distributions of all possible marginals are specified, the symplectic form can be recovered.
-
-In quantum mechanics, the geometry is given by the Born rule
-\begin{equation}
-	p(\phi|\psi) = \frac{\< \phi | \psi \>\< \psi | \phi \>}{\< \psi | \psi \>\< \phi | \phi \>}.
-\end{equation}
-Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, mathematically $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has von Neumann entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi|\psi)}}{2},\frac{1 - \sqrt{p(\phi|\psi)}}{2} \right)$ where $I(x,y) = -x \log x -y \log y$ is the Shannon entropy.
-Therefore the entropy is an invertible function of $p(\phi|\psi)$.
-This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the state space.
 
@@
 \end{enumerate}
+\fi
 
 \section{Proposed research program}
+
+The proposed work starts from the observation that the space of statistical ensembles of both classical mechanics have a lot in common.
+They are both convex subsets of a locally convex topological vector spaces whose metric is induced by the entropy.
+More specifically, by the Jensen-Shannon divergence.
+This is in contrast of the different structure of classical and quantum mechanics, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces respectively.
+Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.
+Born rule) of quantum mechanics can be indeed recovered by the entropy.
+This makes sense as the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
+
+The goal is to find minimal requirements for an entropic structure that is common to all physical theories in such a way that results from classical and quantum theories, in areas like mechanics and information theory, can be generalized.
+In particular, it needs to be understood the interaction between the four mathematical structure required by a space of ensembles: the topology, the convex structure, the proposed entropic structure and the Lie algebra.
+Moreover, it is crucial that these premises are well founded on physical requirements, so that the generalization is of value to physics.
+
 \subsection{Intellectual merit}
 
-The goal is to find a very limited set of axioms to define the entropic structure of a physical theory that can be justified from physical considerations, show that this set of axioms imposes a geometric structure on a generic space of statistical ensembles and that it recovers the standard geometric structures of classical and quantum mechanics.
+The goal of this proposal is to develop a common axiomatic basis for the information theoretic structure underlying to all physical theory, classical and quantum.
+The immediate aim is to generalize results of both classical and quantum information theory to a common more abstract theory that can also handle the infinite dimensional case.
+Longer term, we aim to use the axioms developed as part of a larger effort to give unified base theory for physical theories and clean up some mathematical issues.
+
+While the geometric structure of both classical and quantum mechanics are rather different, their space of ensembles have a rather similar structure: they consist of a bounded convex subset of a metrizable and second countable locally convex topological vector space, where the metric can be derived from the entropy (e.g.
+trough the square root of the Jensen-Shannon divergence).
+There are, therefore, three common structures at play: a topology (which physically represents experimental verifiability), a convex structure (which physically represents the ability to make statistical mixtures) and an entropic structure (which physically represents the variability of the elements within each ensemble).
+While the axioms for the topology and the convex structure are readily available, to our knowledge, the correct axioms for the entropic structure still needs to be identified.
+Moreover, the interplay between the topology, the convex structure and the entropic structure opens additional mathematical questions which have not been investigated.
+
+As a guide to the reader, let us recall how ensembles are defined in both classical and quantum mechanics.
+For a discrete classical system, the state space $X$ is a countable set with the discrete topology, the space of ensembles consists of all probability measure over the Borel algebra of $X$ and the entropy is given by $S(p_i) = - \sum_i p_i \log p_i$.
+Note that all logs here are assumed base 2.
+In classical mechanics, instead, the state space $X$ is a symplectic manifold, which is therefore equipped with the  Liouville measure $\mu(U) = \int_U \omega^n$, where $\omega$ is the symplectic form.
+The space of statistical ensembles is the space of all probability measures over the Borel algebra of $X$ that are absolutely continuous with respect to the Liouville measure.
+For these measures, the Radon-Nikodym derivative $\rho = \frac{dp}{d\mu}$ exists and the entropy is defined as $S(\rho) = - \int \rho \log \rho d\mu$.
+In particular, for a uniform distribution with support $U$, the entropy reduces to $S(\rho_U) = \log \mu(U)$.
+In quantum mechanics, the state space is given by the projective space $P(\mathcal{H})$ of a complex Hilbert space $\mathcal{H}$, the space of ensembles is given by the density operators (i.e.
+positive semi-definite, self-adjoint trace-one operators) and the entropy is given by $S(\rho) = -\tr(\rho \log \rho)$.
+
+Note that in all three cases the definition of the entropy is tightly bound to the definition of the space, therefore the biggest problem is understanding how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each cases.
+Fortunately, in our preliminary work we already have found the solution.
+In all cases, the entropy presents an upper bound to its increase under convex combinations: $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$ where $I(p) = -p \log p - (1-p) \log (1-p)$.
+This maximum increase is given precisely when the two ensembles are orthogonal with respect to the inner product of the respective spaces.
+In the classical case, this means that the measures have disjoint support and in the quantum case that the eigenspaces of non-zero eigenvalues are orthogonal.
+In other words, characterizing the interaction of the entropy function with the topology and the convex structure is enough to characterize the entropic structure.
+We will see that, in fact, the form of $I(p)$, the Shannon entropy formula, can be recovered.
+
+Another important insight from our preliminary work is that the entropy is not only responsible for the geometry of the ensemble space, but it is also intimately related to the geometry of the state space of both classical and quantum mechanics.
+That is, the entropic structure of the ensemble space is equivalent to the geometric structure of the state space.
+As we saw in classical mechanics, the symplectic form gives us the Liouville measure which in turn is used to define the entropy.
+Therefore, the entropy of ensemble is defined from the geometry.
+However, note that the entropy for a uniform distribution $\rho_U$ over a set $U$ is given by $S(\rho_U) = \log \mu(U)$.
+Therefore, if we are given the entropy for all probability measures, we can recover the Liouville measure by setting $\mu(U) = 2^{S(\rho_U)}$.
+While this does not recover the symplectic form directly, note that the symplectic form can also be used to define the entropy of the marginal distributions along each degree of freedom.
+Around a point, in fact, we can define Darboux coordinates and define a joint probability distribution that is the product of independent marginals along each pair.
+One finds a similar relationships between the entropy of uniform marginal distributions and the integral of the symplectic form over the support of the marginal.
+
+In quantum mechanics, the geometry is given by the Born rule
+\begin{equation}
+	p(\phi|\psi) = \frac{\< \phi | \psi \>\< \psi | \phi \>}{\< \psi | \psi \>\< \phi | \phi \>}.
+\end{equation}
+Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, mathematically $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has von Neumann entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi|\psi)}}{2} \right)$ where $I$, as before, is the Shannon entropy for two elements.
+As before, the entropy is defined in terms of the inner product.
+However, the relationship between $S(\rho_{\psi\phi})$ and $p(\phi|\psi)$ is invertible.
+This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the state space.
+
+These insights from our preliminary work shows that the entropic structure plays a crucial role in all physical theories and merits to be studied.
+Moreover, while typically information geometry is casted in finite dimensional spaces, there is nothing preventing its generalization to infinite dimensions.
+If the entropic structure, in fact, imposes a distance function whose open balls are convex, then this provides us with a locally convex topology which is exactly what is needed to define calculus in infinite dimensions.
+Moreover, the Lie algebras of both classical (i.e.
+Poisson brackets) and quantum (i.e.
+commutators) have a special relationship with the geometric and entropic structure, which may provide us with a template for generalizes those structure as well.
+
+Given these very promising preliminary findings, the goal is to find a minimal set of axioms to characterize the entropic structure.
+Minimal in the sense that, in the spirit of analogous work on the foundations of set theory or reverse mathematics, the axioms can be shown to be satisfied by the different theories (which now become model of the larger theory) but also allow to rederive the entropy formula for each specific theory.
+Let us, then, go through the current state of our preliminary work, and show the type of questions that the overall approach touches.
+
+%The goal is to find a very limited set of axioms to define the entropic structure of a physical theory that can be justified from physical considerations, show that this set of axioms imposes a geometric structure on a generic space of statistical ensembles and that it recovers the standard geometric structures of classical and quantum mechanics.
 Furthermore, the same structure should be connected to topological and measure theoretic structures used in physics.
@@
 
-We want to stress that finding the correct starting point is not just a mathematical issue, but also a physics problem.
+%We want to stress that finding the correct starting point is not just a mathematical issue, but also a physics problem.
 Let us explore this problem through an example.
@@
 
-Our preliminary work convinced us that the notion of statistical ensemble (i.e.~the collection of all outputs of a preparation procedure) seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
-In terms of the math, the structure of statistical ensembles of both classical and quantum mechanics can be made to look very similar.
+The central focus is the intuitive notion of statistical ensemble (i.e.~the collection of all outputs of a preparation procedure), as this seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
+In terms of the math, as we have seen, the structure of statistical ensembles of both classical and quantum mechanics can be made to look very similar.
 From the experimental side, statistical ensembles are what we produce and measure experimentally.
@@
 Therefore, it makes sense to require that any physical theory must \emph{at least} posit a space of ensembles.
-Our preliminary work shows that there are three, possibly four, distinct sets of requirements for an ensemble space $\Ens$, the set of ensembles given by a physical theory.
-The work proposed here focuses on further development of the third.
+This provides us with a strong conceptual starting point and it is properties of ensembles that needs to be specified through axioms.
 
-The first requirement is that of experimental verifiability of the theory.
-This has been thoroughly investigated in our previous work that established a precise correspondence between experimental verifiable statements (i.e.~statements associated with an experimental test that terminates successfully in finite time if and only if the statement is true) and the open set of a ``natural'' topology, and between theoretical statements (i.e.~statement associated with an experimental test without termination guarantees) and Borel sets~\cite{aop-book,aop-topExpDisting}.
+The first axiom establishes that an ensemble space $\Ens$ is a $\mathsf{T}_0$ second-countable topological space.
+This is based on our previous work, that established a link between the experimental verifiability of the theory and its topology.
+The open sets of the topology correspond to experimentally verifiable statements (i.e.~statements associated with an experimental test that terminates successfully in finite time if and only if the statement is true) while Borel sets correspond to statements associated with an experimental test without termination guarantees~\cite{aop-book,aop-topExpDisting}.
 Real and integer numbers, for example, are recovered from an idealized model representing references of a reference system.
 The key insight is that the set-theoretic operations represent a two-valued logic system (i.e.~truth or falsehood of statements) while the topology keeps track of a three-valued logic system (i.e.~interior, exterior and boundary represent respectively successful termination, unsuccessful termination and non-termination).
-This work gives us a solid foundation on which to build, and guarantees that an ensemble space $\Ens$ is, at least, a $\mathsf{T}_0$ second-countable topological space.
+The topology, therefore, plays an important role.
 
-The second requirement is the ability to perform statistical mixtures.
+The second axiom imposes a convex structure which represent the ability to perform statistical mixtures.
 Given two ensembles $\ens[a], \ens[b] \in \Ens$, we can always choose a probability $p \in [0,1]$ and select $\ens[a]$ $p$ percent of the times and $\ens[b]$ $\bar{p} = 1-p$ percent of the times.
@@
 On physical grounds, the mixing operation must be continuous in the natural topology defined by experimental verifiability, much like addition and scalar multiplication are continuous in a topological vector space, leading to the idea of a topological convex space.
-We haven't found such a structure already studied in the literature.
-While it would be extremely useful to us to see what can be generalized from topological vector spaces to topological convex spaces, we do not believe that domain knowledge in physics is specifically valuable here and leave this work to others, as we believe that our expertise in the mathematical foundations of physical theories is better spent in continuing to find other well defined starting points.
 
-The third requirement, central to the proposed work, is characterizing the variability of elements within a statistical ensemble.
+Surprisingly, topological convex spaces do not seem to have been studied, and this structure alone presents interesting sets of questions.
+For example, it is established that a convex space embeds into a vector space if and only if it is cancellative.
+That is, if $p \ens[a] + \bar{p} \ens = p \ens[b] + \bar{p} \ens$ for some $p \in (0,1)$ implies $\ens[a] = \ens[b]$.
+The question now is what are necessary and sufficient conditions for a topological convex space to embed continuously into a topological vector space?
+Can we have, for example, an Hausdorff second countable topological convex space that is not a continuous embedding of a topological vector space?
+Topological vector space present a number of nice results which may or may not hold in topological convex spaces.
+For example, any $\mathsf{T}_1$ second countable topological vector spaces is metrizable.
+Does that same result hold for topological convex spaces?
+Another interesting aspect is the notion of closure.
+In a topological vector space, the topological closure of a convex set is convex.
+We found that this is true in topological convex spaces as well.
+Additionally, there is another type of closure that may be more interesting physically.
+We can define the $\sigma$-hull of a set $A$ the set of possible infinite convex combinations, which can also be shown to be a convex set.
+This will be a superset of the hull of $A$, because it will include all finite convex combinations, and it will be a subset of the closed hull of $A$, as all infinite convex combinations are topological limits.
+However, it can be shown that, in some case, the hull, the $\sigma$-hull and the closed hull are all different.
+From the physics side, the $\sigma$-hull may be a more physically meaningful closure as it represents the true limit process of statistical mixing, the operation the axiom is characterizing.
+Therefore, there are already interesting topics to study even without imposing the entropic structure.
+
+The third axiom, central to the proposed work, imposes an entropy function that quantifies the variability of elements within a statistical ensemble.
 Though we imagine ensembles to be a collection of instances, these instances are never formalized as they need to be left unspecified to leave the theory general.
@@
 
-The hardest conceptual problem we found is that the entropy is defined differently in each space.
-In a classical discrete space, the entropy is $-\sum p_i \log p_i$; in classical statistical mechanics it is $- \int \rho \log \rho d\mu$ where $\mu$ is the Liouville measure; in quantum mechanics it is $-\tr(\rho \log \rho)$.
-How can we give a single definition for these very different expressions?
-Note that the expressions are different precisely because they are in terms of the specific representation of ensembles in the specific theory.
-Our approach is to simply characterize the entropy by its properties with respect to the topology and the convex structure.
-In our preliminary work, for example, the axiom defining entropy $S : \Ens \to \mathbb{R}$ currently has the following properties:
+In our preliminary work, for example, the axiom equips each ensemble space with an entropy function $S : \Ens \to \mathbb{R}$ currently has the following properties:
 \begin{itemize}
@@
 	\item Strict concavity: $S(p\ens[a] + \bar{p} \ens[b]) \geq p S(\ens[a]) + \bar{p} S(\ens[b])$ with the equality holding if and only if $\ens[a] = \ens[b]$
-	\item Upper variability bound: there exists a universal function $I(p_1, p_2)$ (i.e.~the same for all ensemble spaces) such that $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p, \bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])$; if the equality holds, $\ens[a]$ and $\ens[b]$ are orthogonal, noted $\ens[a] \ortho \ens[b]$
+	\item Upper variability bound: there exists a universal function $I : [0,1] \to \mathbb{R}$ (i.e.~the same for all ensemble spaces) such that $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$; if the equality holds, $\ens[a]$ and $\ens[b]$ are orthogonal, noted $\ens[a] \ortho \ens[b]$
 	\item Mixtures preserve orthogonality: $\ens[a] \ortho \ens[b]$ and $\ens[a] \ortho \ens[c]$ if and only if $\ens[a] \ortho p \ens[b] + \bar{p} \ens[c]$ for any $p \in (0,1)$
 \end{itemize}
-These properties are justified by the physics.
+These properties can be justified by the physics.
 For example, if we are performing a statistical mixture between two ensembles, the average variability cannot decrease, which justifies the concavity.
-These relatively simple properties allow us, for example, to show that $I(p,\bar{p})$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy~\cite{Shannon} $ - \left(p \log p + \bar{p} \log \bar{p}\right)$.
+The goal is to have a full justification for each property to make sure that we are not leaving out any possibly interesting physical theories.
+Similarly to one does in foundations of mathematics, we typically construct different models of the various formulations of the axioms to check that each property is truly independent from the other, and that is physically required.
+
+Though we do not believe the axiom lists all necessary properties of the entropic structure, it does list enough of them to already prove interesting results.
+For example, we can show that $I(p)$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy~\cite{Shannon} $I(p)= - p \log p - \bar{p} \log \bar{p}$.
 The proof is general, meaning that if we are right that those axioms must hold for any physical theory, then the Shannon entropy is of truly general applicability.
 Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is an equality and can be used to calculate the entropy from that decomposition.
-This allows us to recover the expressions for the entropy in the three different cases, classical discete, classical continuous, and quantum.
+This will allows us to recover the expressions for the entropy in the three different cases, classical discete, classical continuous, and quantum.
+Similarly, we will be able to show that, from this structure alone, the geometric structure of both classical and quantum mechanics can be recovered.
 
-Our preliminary work shows that an entropy satisfying those properties forces the convex space to be cancellative, meaning that it embeds into a vector space.
-That is, if we pick an internal point $\ens \in \Ens$, we can define a vector space $V_{\ens}$ such that there exists an embedding $\iota : \Ens \hookrightarrow  V_{\ens}$ such that $\iota(p \ens[a] + \bar{p} \ens[b]) = p \iota(\ens[a]) + \bar{p} \iota(\ens[b])$ and $\iota(\ens) = 0$.
-If we pick two different internal points $\ens[a],\ens[b] \in \Ens$, we can show that there is an invertible affine map between $V_{\ens[a]}$ and $V_{\ens[b]}$.
-That is, the ensemble space should be more properly thought as embedding into an affine space.
-We can also show that it is ``directionally bounded,'' meaning its intersection with any affine line is a finite interval.
-It is an open question whether the currently defined properties are enough to prove that the embedding is continuous in a topological vector space, and whether the ensemble space is a bounded subset.
+Our preliminary work also shows that an entropy satisfying those properties forces the convex space to be cancellative, meaning that it embeds into a vector space.
+That is, a convex space that does not embed into a vector space cannot posses an entropic structure.
+We can, in fact, produce a model of that case and see exactly why the entropic structure fails.
+We can also show ensemble spaces must be ``directionally bounded,'' meaning its intersection with any affine line is a finite interval.
+It is an open question as to whether the ensemble must be a bounded set.
+We can also show that some models of GPTs (i.e.
+Generalized Probabilistic Theories) are inconsistent with an entropic structure and therefore cannot represent a physically meaningful space of ensembles.
+In other words, the entropic structure constrain the type of spaces possible in surprising and interesting ways.
 
-We can also show that the entropy can be used to define geometric structures directly on the ensemble space.
+Another line of inquiry is the relationship between the entropic structure and a metric ensemble space.
 For example, we can define the mixing entropy as $MS(\ens[a], \ens[b]) = S\left(\frac{1}{2}\ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2} S(\ens[a]) + \frac{1}{2} S(\ens[b])\right)$.
@@
 While the JSD is the square of a distance function in both classical and quantum information theory, our current axioms do not guarantee $MS$ to have this property in general.
-It is an open question whether enough physical constraints exist for that to be the case.
+An interesting question, then, is whether enough physical constraints exist for that to be the case.
 The $MS$ can be used to create open balls and show that the topology must be Hausdorff.
-An interesting conjecture is whether a Hausdorff second-countable topological convex space is metrizable, given that this is true for topological vector spaces.
-The current axioms do not guarantee the open balls to be convex.
-If we were to find sufficient physically motivated conditions that imply that the entropic open balls are convex and generate the topology, as we suspect should be the case, the topology would be locally convex and therefore the entropic structure also may give us enough structure to define calculus.
+It is not yet clear whether this topology is metrizable and whether it must coincide with the topology of the ensemble space.
 
-If we assume we are in finite dimensions and that the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
+A related line of questions relates the entropic structure to elements of differential geometry.
+First, is the relationship to differentiability.
+We are aware that the current axioms do not guarantee the entropic open balls to be convex, even though they are in both the classical and quantum case.
+We suspect that to be the case in general, even though we haven't yet found a physically satisfying premise.
+If the convexity of the open ball is established, then the topology generated by entropic open balls is locally convex, which give us enough structure to define calculus in infinite dimensional spaces.
+That is, the entropic structure would be responsible for the ability to define derivatives.
+
+In the final dimensional case, and if the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
 Let $V_{\ens[a]}$ be the embedding vector space with $\ens[a] \in \Ens$ as origin.
@@
 In classical and quantum mechanics it recovers the Fisher information metric.
-The ultimate goal would be to show that, in the general case, the embedding is continuous into a locally convex topological vector space so that this construction can be extended to the infinite-dimensional case.
-This would essentially show that the entropy provides, in general, the geometric structure of the state space of any physical theory.
+The immediate goal, then, is to see how much of information geometry can be generalized to this case.
+If a locally convex topology can be established, the further goal is to extend this to the infinite dimensional case.
 
-The fourth requirement, which is left for future work, is the generalization of the Lie brackets for a generic ensemble space, given by the Poisson brackets in the classical case and the commutators in the quantum case.
-These describe the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged, which means, as outlined before, leaving the entropic structure unchanged.
+Another goal is to recover the geometric structure of the state spaces of both classical and quantum mechanics from the entropic structure, to formally show the equivalence.
+For quantum mechanics, one can simply define the Born rule from the entropy as we have seen before.
+For classical mechanics, the situation is a bit more complex as we need to recover first the Liouville measure.
+This can be done by noting two things: for a uniform distribution with support $U \subseteq X$, $\mu(U) = 2^{S(\rho_U)}$ and the uniform distribution over $U$ has the highest entropy of all distributions with support $U$.
+Given a set $A \subseteq \Ens$ of ensemble, we can then define the state capacity as supremum of the exponential of the hull of $A$, which physically means the highest entropy achievable with a statistical mixture of elements of $A$.
+This set function has nice properties: it is non-negative, monotonic, subadditive in general and additive on orthogonal sets of ensembles.
+Note that if $A_U \subseteq \Ens$ is the subspace of all probability measures with support $U \subseteq X$, $A_U$ is orthogonal to $A_V$ if and only if $U$ and $V$ are disjoint.
+This means that the state capacity restricted to the lattice of those subspaces is an additive measure.
+Interestingly enough, one can follow the same construction in the quantum case, and the measure will return the dimensionality of the subspace.
+In the areas of non-additive measure, there are more interesting questions and one can create a non-additive version of probability, which may point to a non-additive version of probability theory.
+We have already looked at this problem, and, unfortunately, the standard techniques that have been developed in that area are not sufficient.
+
+A possible fourth axiom maybe require to conclude all the requirements from physical theories.
+In both classical and quantum mechanics, there a fundamental Lie algebra on the statistical variables.
+In classical mechanics, this is captured by the Poisson bracket while in quantum mechanics is captured by the commutator.
+Since the set of the probability measures defined on an odd dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the above structure is not sufficient.
+However, there is an intimate relationship between this Lie algebraic structure and the entropic structure since, in both cases, they describe the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged, which means, as outlined before, leaving the entropic structure unchanged.
 That is, it describes the group of deterministic and reversible motions, those that leave the entropy unchanged~\cite{aop-HamConsInfoEnt,aop-HamQuantInfo,aop-book}.
-Given the tight connection between the entropic structure and the Lie-algebraic structure, it is unclear whether this will be recovered automatically once the entropic structure is correctly axiomatized, or whether it will require additional axioms.
-For this reason, we postpone the study of this aspect.
+Part of the investigation for the correct specification of the entropic structure, then, will need to consider its impact on the definition of the Lie algebraic structure.
+For example, in our reverse physics approach to classical mechanics we were able to show that a symplectic structure is necessary and sufficient to define an entropy that is invariant under frame transformations.
+If such requirement could be specified in general, the principle of relativity could be shown to be equivalent to this structure.
 
-Having given an overview of the specifics of the problems, the proposed work consists of first finalizing and publishing some of the preliminary results, which include the equivalence of geometric and entropic structures in physical theories, the ability to recover the Shannon entropy in the generalized case and the three different entropies in the discrete/continuous/quantum cases.
-At the same time, we will explore whether the axiomatics can be expanded without sacrificing potentially interesting physical cases in order to recover potentially useful properties, such as the metrizability of the space and local convexity of the topology generated by the entropy.
+As we have seen, the program is rich and connects to many areas of physics, information theory and mathematics.
+To not get lost, let us reiterate the primary goals.
+We want to mathematically formalize the intuition that specifying the geometry in physics is equivalent to specify the entropy.
+To do so, we need to give a precise, physically motivated, axiomatic definition of what an entropic structure is for a general theory.
+Then show that this characterized the geometry of the general space in a way that recovers the geometry of classical and quantum mechanics in the specific case.
+While doing that, we will gain insight on how to handle topological issues properly, and hopefully show that ensemble spaces must embed into a second countable metrizable locally convex vector space, so that notion of differentiability can be extended to the infinite dimensional case.
+We will also keep an eye open to connection to measure theory and Lie algebra, so that more features of physical theories can be generalize in the future.
 

### 2025-10-20 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Final touches

@@
 The immediate aim is to generalize results of both classical and quantum information theory to a common more abstract theory that can also handle the infinite dimensional case.
-Longer term, we aim to use the axioms developed as part of a larger effort to give unified base theory for physical theories and clean up some mathematical issues.
+Longer term, we aim to use the axioms developed as part of a larger effort to develop a unified base theory for physical theories and clean up some mathematical issues.
 
@@
 In the classical case, this means that the measures have disjoint support and in the quantum case that the eigenspaces of non-zero eigenvalues are orthogonal.
-In other words, characterizing the interaction of the entropy function with the topology and the convex structure is enough to characterize the entropic structure.
-We will see that, in fact, the form of $I(p)$, the Shannon entropy formula, can be recovered.
+In other words, characterizing the interaction of the entropy function with the topology and the convex structure is enough to characterize the entropic structure: the form of $I(p)$, the Shannon entropy formula, and, since in both classical and quantum mechanics ensembles can always be decomposed into orthogonal component, the different entropy formulas can be recovered in each case.
 
@@
 While this does not recover the symplectic form directly, note that the symplectic form can also be used to define the entropy of the marginal distributions along each degree of freedom.
-Around a point, in fact, we can define Darboux coordinates and define a joint probability distribution that is the product of independent marginals along each pair.
+Around a point, in fact, we can define Darboux coordinates and define joint probability distributions that are the product of independent marginals along each pair of position and momentum.
 One finds a similar relationships between the entropy of uniform marginal distributions and the integral of the symplectic form over the support of the marginal.
@@
 However, the relationship between $S(\rho_{\psi\phi})$ and $p(\phi|\psi)$ is invertible.
-This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the state space.
+This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the geometry of the state space.
 
 These insights from our preliminary work shows that the entropic structure plays a crucial role in all physical theories and merits to be studied.
-Moreover, while typically information geometry is casted in finite dimensional spaces, there is nothing preventing its generalization to infinite dimensions.
+Moreover, while typically information geometry is cast in finite dimensional spaces, there is nothing preventing its generalization to infinite dimensions.
 If the entropic structure, in fact, imposes a distance function whose open balls are convex, then this provides us with a locally convex topology which is exactly what is needed to define calculus in infinite dimensions.
@@
 The central focus is the intuitive notion of statistical ensemble (i.e.~the collection of all outputs of a preparation procedure), as this seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
-In terms of the math, as we have seen, the structure of statistical ensembles of both classical and quantum mechanics can be made to look very similar.
+In terms of the math, as we have seen, the spaces of statistical ensembles of classical and quantum mechanics can be made to look very similar.
 From the experimental side, statistical ensembles are what we produce and measure experimentally.
@@
 
-The second axiom imposes a convex structure which represent the ability to perform statistical mixtures.
+The second axiom imposes a convex structure which represents the ability to perform statistical mixtures.
 Given two ensembles $\ens[a], \ens[b] \in \Ens$, we can always choose a probability $p \in [0,1]$ and select $\ens[a]$ $p$ percent of the times and $\ens[b]$ $\bar{p} = 1-p$ percent of the times.
@@
 Can we have, for example, an Hausdorff second countable topological convex space that is not a continuous embedding of a topological vector space?
-Topological vector space present a number of nice results which may or may not hold in topological convex spaces.
+Topological vector spaces present a number of nice results which may or may not hold in topological convex spaces.
 For example, any $\mathsf{T}_1$ second countable topological vector spaces is metrizable.
@@
 For example, if we are performing a statistical mixture between two ensembles, the average variability cannot decrease, which justifies the concavity.
-The goal is to have a full justification for each property to make sure that we are not leaving out any possibly interesting physical theories.
+The goal is to have a full justification for each property to make sure that we are not leaving out any possibly interesting physical theory.
 Similarly to one does in foundations of mathematics, we typically construct different models of the various formulations of the axioms to check that each property is truly independent from the other, and that is physically required.
@@
 For example, we can show that $I(p)$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy~\cite{Shannon} $I(p)= - p \log p - \bar{p} \log \bar{p}$.
-The proof is general, meaning that if we are right that those axioms must hold for any physical theory, then the Shannon entropy is of truly general applicability.
+The proof is general in the sense that it does not assume whether the specific model is classical, quantum or something else entirely.
 Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is an equality and can be used to calculate the entropy from that decomposition.
@@
 
-Another line of inquiry is the relationship between the entropic structure and a metric ensemble space.
+Another line of inquiry is the relationship between the entropic structure and a metric over the ensemble space.
 For example, we can define the mixing entropy as $MS(\ens[a], \ens[b]) = S\left(\frac{1}{2}\ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2} S(\ens[a]) + \frac{1}{2} S(\ens[b])\right)$.
@@
 
-A related line of questions relates the entropic structure to elements of differential geometry.
-First, is the relationship to differentiability.
+A related line of questions relates the entropic structure to elements of differential geometry and calculus of variation.
 We are aware that the current axioms do not guarantee the entropic open balls to be convex, even though they are in both the classical and quantum case.
@@
 
-In the final dimensional case, and if the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
+In the finite dimensional case, and if the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
 Let $V_{\ens[a]}$ be the embedding vector space with $\ens[a] \in \Ens$ as origin.
@@
 This set function has nice properties: it is non-negative, monotonic, subadditive in general and additive on orthogonal sets of ensembles.
-Note that if $A_U \subseteq \Ens$ is the subspace of all probability measures with support $U \subseteq X$, $A_U$ is orthogonal to $A_V$ if and only if $U$ and $V$ are disjoint.
+Note that if $A_U \subseteq \Ens$ is the subspace of all probability measures with support $U \subseteq X$, the state capacity of $A_U$ recovers the Liouville measure of $I$.
+Moreover, $A_U$ is orthogonal to $A_V$ if and only if $U$ and $V$ are disjoint.
 This means that the state capacity restricted to the lattice of those subspaces is an additive measure.
 Interestingly enough, one can follow the same construction in the quantum case, and the measure will return the dimensionality of the subspace.
-In the areas of non-additive measure, there are more interesting questions and one can create a non-additive version of probability, which may point to a non-additive version of probability theory.
+Concerning non-additive measure, there are more interesting questions as one can create a non-additive version of probability, which may point to a non-additive version of probability theory.
 We have already looked at this problem, and, unfortunately, the standard techniques that have been developed in that area are not sufficient.
 
-A possible fourth axiom maybe require to conclude all the requirements from physical theories.
-In both classical and quantum mechanics, there a fundamental Lie algebra on the statistical variables.
+A possible fourth axiom may be required to conclude all the requirements for physical theories.
+In both classical and quantum mechanics, there is a Lie algebra that characterizes the relationships between generators of time evolution.
 In classical mechanics, this is captured by the Poisson bracket while in quantum mechanics is captured by the commutator.
-Since the set of the probability measures defined on an odd dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the above structure is not sufficient.
-However, there is an intimate relationship between this Lie algebraic structure and the entropic structure since, in both cases, they describe the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged, which means, as outlined before, leaving the entropic structure unchanged.
+Since the set of the probability measures defined on an odd dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the previous three axioms are not sufficient.
+However, there is an intimate relationship between this Lie algebraic structure and the entropic structure since, in both cases, they describe the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged, which means, as outlined before, they must leave the entropic structure unchanged.
 That is, it describes the group of deterministic and reversible motions, those that leave the entropy unchanged~\cite{aop-HamConsInfoEnt,aop-HamQuantInfo,aop-book}.
@@
 For example, in our reverse physics approach to classical mechanics we were able to show that a symplectic structure is necessary and sufficient to define an entropy that is invariant under frame transformations.
-If such requirement could be specified in general, the principle of relativity could be shown to be equivalent to this structure.
+If such requirement could be specified in general, the principle of relativity at equal time could be shown to be equivalent to this structure.
 

### 2025-10-20 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Tasks for student

@@
 
+Depending on the interests and skill of the graduate student, there are different ways to contribute.
+One possibility is to analyze what constructions and theorem from topological vector spaces can be generalized to topological convex spaces.
+This will help us better understand the relationship between the two, and possibly give insights in what properties the entropic structure may impose to bridge the gap.
+Another possibility is to see what results from classical and quantum information theory can be generalized to the entropic structure alone.
+Yet another possibility, since the JSD is the square of a distance function, finding sufficient conditions for a pseudo-distance to be the square of a distance my clarify what properties the entropic structure may impose.
+
 Notes:

### 2025-10-24 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Updates mostly to nontechnical sections

--- 2025-10-20 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
@@
 \begin{center}
-	\textbf{A generalized axiomatic basis for entropic structures in physical theories} \\
-	Christine A. Aidala, University of Michigan
+	\textbf{An axiomatic basis for entropic structures in physical theories} \\
+	Christine A. Aidala (PI) and Alejandro Uribe (Co-PI), University of Michigan
 \end{center}
@@
 The proposed work starts from the observation that the space of statistical ensembles of both classical mechanics have a lot in common.
-They are both convex subsets of a locally convex topological vector spaces whose metric is induced by the entropy.
+They are both convex subsets of a locally convex topological vector space whose metric is induced by the entropy.
 More specifically, by the Jensen-Shannon divergence.
-This is in contrast of the different structure of classical and quantum mechanics, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces respectively.
-Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.
-Born rule) of quantum mechanics can be indeed recovered by the entropy.
+This is in contrast to the different structures of classical and quantum mechanics, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces, respectively.
+Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can indeed be recovered by the entropy.
 This makes sense as the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
 
-The goal is to find minimal requirements for an entropic structure that is common to all physical theories in such a way that results from classical and quantum theories, in areas like mechanics and information theory, can be generalized.
-In particular, it needs to be understood the interaction between the four mathematical structure required by a space of ensembles: the topology, the convex structure, the proposed entropic structure and the Lie algebra.
+The goal is to find minimal requirements for an entropic structure that is common to all physical theories in such a way that known results from classical and quantum theories, in areas like mechanics and information theory, can be generalized.
+In particular, the relationships among the four mathematical structures required by a space of ensembles---the topology, the convex structure, the proposed entropic structure and the Lie algebra---need to be understood.
 Moreover, it is crucial that these premises are well founded on physical requirements, so that the generalization is of value to physics.
@@
 \noindent
-\textbf{Relationship to other current funding:} The PI's current funding from the NSF Division of Physics is for work on LHCb experiment at CERN, and from the NSF AccelNet program is to increase collaboration on quantum chromodynamics (QCD) research between the U.S.~and Latin America.
+\textbf{Relationship to other current funding:} PI Aidala's current funding from the NSF Division of Physics is for work on LHCb experiment at CERN, and from the NSF AccelNet program is to increase collaboration on quantum chromodynamics (QCD) research between the U.S.~and Latin America.
 The PI's funding from the Department of Energy is for work at the Relativistic Heavy Ion Collider (RHIC) on the PHENIX and sPHENIX experiments.
@@
 Her grant from the John Templeton Foundation, titled ``Why classical probability and classical information theory are incompatible with quantum mechanics and quantum contextuality,'' supports the development of novel non-additive measures (set functions) suitable for quantifying the number of states in quantum mechanics~\cite{aop-nonaddmeas}.
-There is no direct overlap with the activities proposed here.
+There is no direct overlap with the activities proposed here, and the end date of the grant is moreover before the start date of the work proposed here.
 \\
+
+Co-PI Uribe does not have any current funding. \\
 
@@
 
+TODO: Need to add 1-2 sentences on other commitments of Alejandro as Co-PI?
+
 \section{Personnel}
-The proposed research will be performed by PI Christine Aidala, her collaborator Gabriele Carcassi, and Ph.D.~student Tobias Thrien, all part of the Physics Department at the University of Michigan.
-See also the section on ``Qualifications of the PI and Research Team'' below.
+The proposed research will be performed by PI Christine Aidala (U-M Physics), Co-PI Alejandro Uribe (U-M Mathematics), staff researcher Gabriele Carcassi (U-M Physics), and a Ph.D.~student to be recruited from the Applied and Interdisciplinary Mathematics (AIM), Physics, or Mathematics Ph.D.~program.
+See also the section on ``Qualifications of the Research Team'' below.
 \\
-TODO: Add Alejandro if he agrees.
 
-Aidala as PI will be responsible for coordinating all of the work, and she will serve as the academic advisor for Thrien.
-She will identify suitable journals for publication and conferences at which to disseminate the results, and oversee the direction of the work.
+Aidala as PI will be responsible for coordinating all of the work.
+She and/or Uribe will serve as the formal academic (co-)advisor for the Ph.D.~student.
+Students in the AIM program are required to have a primary advisor in Mathematics and a co-advisor in another Department.
+Aidala and Uribe will identify suitable journals for publication and conferences at which to disseminate the results, and oversee the direction of the work.
+Uribe will
 
@@
 He is (currently part-time) staff in the U-M Department of Physics, a Research Area Specialist Senior, with his salary fully supported through grants.
-He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until Thrien is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.
-With training in engineering and mathematics, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
+He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until the graduate student is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.
+With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
 
-Thrien entered the U-M Physics Ph.D.~program in Fall 2024.
-While Carcassi will be the primary individual responsible for ``outward-facing'' technical work, Thrien will focus primarily on inward-facing technical work.
-\\
-TODO: Be more specific about Thrien's work, per reviews from last year.
+Depending on the interests, background, and skills of the Ph.D.~student, there will be different ways for the student to contribute.
+One possibility is to analyze what constructions and theorems from topological vector spaces can be generalized to topological convex spaces.
+This will help us better understand the relationship between the two, and possibly give insights into what properties the entropic structure may impose to bridge the gap.
+Another possibility is to see what results from classical and quantum information theory can be generalized to the entropic structure alone.
+Yet another possibility, since the JSD is the square of a distance function, is to identify sufficient conditions for a pseudo-distance to be the square of a distance, which may clarify what properties the entropic structure imposes.
 
@@
 
-\section{Qualifications of the PI and research team}
+\section{Qualifications of the research team}
 PI Christine Aidala is a highly recognized leader within physics, having received an NSF CAREER Award, a Sloan Fellowship, and a Presidential Early Career Award for Scientists and Engineers.
@@
 
-TODO: Add info for Alejandro if agrees.
+TODO: Add info for Alejandro.
 
@@
 He and Aidala gave a joint public talk for the Michigan State University Advanced Studies Gateway, and he will give a U-M ``Saturday Morning Physics'' public talk in November 2025.
-His training in engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from theoretical computer science, information theory, systems theory, signal processing, and operations research.
+He was invited and will deliver an overview talk on the AoP project at a special session of the Joint Mathematics Meeting in January 2026.
+[Are all special session talks invited?] His training in computer science and engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from information theory, theoretical computer science, systems theory, signal processing, and operations research.
 His position as U-M Physics Department staff without teaching responsibilities provides him with time and flexibility to dedicate to research related to the AoP project such as the work proposed here.
-\\
+[Some material repetitive--edit.]\\
 
-Together, Aidala and Carcassi [add Alejandro if agrees] are uniquely positioned to advance the mathematical foundations of physics, with Aidala's traditional training in physics and traditional academic position complementing Carcassi's formal training in engineering and his university staff position, which for example gave him the time and flexibility to audit multiple graduate Mathematics courses at U-M.
-For the AoP project, their published work so far has covered a wide range of topics.
+Together, Aidala, Uribe, and Carcassi are uniquely positioned to advance the mathematical foundations of physics, with Aidala's and Uribe's traditional training in physics and mathematics, respectively, complementing Carcassi's formal training in computer science and engineering.
+Carcassi's university staff position has given him the time and flexibility to audit multiple graduate Mathematics and Physics courses at U-M, after having previously audited multiple graduate Physics courses at Stony Brook University.
+Uribe has served as a mathematical consultant to Aidala and Carcassi for the broader Assumptions of Physics project since 2018.
+For the AoP project, Aidala and Carcassi's published work so far has covered a wide range of topics.
 The ``physical mathematics'' approach within their work is an approach to the mathematical foundations of physics that seeks to construct mathematical structures strictly from axioms and definitions that can be rigorously justified from physical requirements~\cite{aop-book}.
-For example, they have shown that experimental distinguishability is mathematically captured by Hausdorff topologies~\cite{aop-book,aop-topExpDisting}, and they have argued that at the smallest space-time scales, at which geometric information becomes in principle experimentally inaccessible, there should still in principle be a regime in which topological structure remains experimentally accessible~\cite{aop-spacetimeStruct}.
-Their ``reverse physics'' approach instead is an approach to the foundations of physics that analyzes known theories to identify those physical principles and assumptions that can be taken as their conceptual foundation~\cite{aop-book,Carcassi:2022bpm}.
+For example, they have shown that experimental distinguishability is mathematically captured by Hausdorff topologies~\cite{aop-book,aop-topExpDisting}, and they have argued that at the smallest space-time scales, at which geometric information becomes experimentally inaccessible, there should still in principle be a regime in which topological structure remains experimentally accessible~\cite{aop-spacetimeStruct}.
+Their ``reverse physics'' approach is instead an approach to the foundations of physics that analyzes known theories to identify those physical principles and assumptions that can be taken as their conceptual foundation~\cite{aop-book,Carcassi:2022bpm}.
 Their reverse physics approach has led for example to a geometric and physical interpretation of the principle of least action~\cite{aop-action} and to a demonstration that the uncertainty principle in quantum mechanics emerges solely from the lower limit on the entropy~\cite{Carcassi:2022bpm}.
@@
 
-Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds, and in 2025 the U-M Rackham Distinguished Graduate Mentor Award.
-She has served as the dissertation chair for 10 graduate students in Physics and Applied Physics who have already completed their doctorates, as co-chair with Uribe for one in Applied and Indisciplinary Mathematics and another in Applied Physics, and is currently the dissertation chair for an additional 4 Ph.D.~students in Physics.
-Aidala and Carcassi have also successfully published with former U-M Ph.D.~students in Mathematics~\cite{aop-spacetimeStruct} and Philosophy~\cite{aop-HamPriv} whose primary dissertation work was with other faculty members.
+Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds, and in 2025 received the U-M Rackham Distinguished Graduate Mentor Award.
+She has served as the dissertation chair for 10 graduate students in Physics and Applied Physics who have already completed their doctorates, as co-chair for another in Applied Physics and as co-chair with Uribe for one in Applied and Indisciplinary Mathematics.
+She is currently the dissertation chair for an additional 4 Ph.D.~students in Physics.
+Aidala and Carcassi have also successfully published with former U-M Ph.D.~students in Mathematics~\cite{aop-topExpDisting} and Philosophy~\cite{aop-HamPriv,aop-unphysHilbert} whose primary dissertation work was with other faculty members.
 In 2020 Aidala received the U-M John Dewey Award for long-term commitment to the education of undergraduate students.
-Between Aidala and Carcassi, they have mentored more than 50 undergraduate students in research since coming to U-M in 2012.
-\\
+Between Aidala and Carcassi, they have mentored more than 50 undergraduate students in research since coming to U-M in 2012, with Carcassi serving as the primary mentor for three undergraduate senior theses related to the AoP project.
+[More on Carcassi's mentoring of undergrads?]\\
 
 TODO: Add info for Alejandro.
+
+Uribe has also served on the dissertation committee of four of Aidala's students thus far.
 
@@
 In more creative outreach work, in April 2024 at U-M he performed an original musical comedy lecture, ``An Engineer in the Foundations of Physics,'' which drew approximately 100 attendees.
+He recently performed a slightly updated version of the comedy lecture at the American Physical Society meeting in Chicago in October 2025.
 

### 2025-10-25 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
To send to Alejandro for him to add his material

@@
 
-The proposed work starts from the observation that the space of statistical ensembles of both classical mechanics have a lot in common.
-They are both convex subsets of a locally convex topological vector space whose metric is induced by the entropy.
-More specifically, by the Jensen-Shannon divergence.
-This is in contrast to the different structures of classical and quantum mechanics, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces, respectively.
+The proposed work starts from the observation that the spaces of statistical ensembles of classical and quantum mechanics have multiple features in common.
+They are both convex subsets of a locally convex topological vector space whose metric is induced by the entropy, or more specifically, by the Jensen-Shannon divergence.
+This is in contrast to the different geometric structures of classical and quantum mechanics, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces, respectively.
 Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can indeed be recovered by the entropy.
-This makes sense as the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
+The states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
 
@@
 
-The goal of this proposal is to develop a common axiomatic basis for the information theoretic structure underlying to all physical theory, classical and quantum.
-The immediate aim is to generalize results of both classical and quantum information theory to a common more abstract theory that can also handle the infinite dimensional case.
-Longer term, we aim to use the axioms developed as part of a larger effort to develop a unified base theory for physical theories and clean up some mathematical issues.
+The goal of this proposal is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum.
+The immediate aim is to generalize results of both classical and quantum information theory to a common, more abstract theory that can also handle the infinite-dimensional case.
+Longer term, we aim to use the axioms developed here as part of a larger effort to develop a unified base theory for physical theories with a clean mathematical foundation.
 
-While the geometric structure of both classical and quantum mechanics are rather different, their space of ensembles have a rather similar structure: they consist of a bounded convex subset of a metrizable and second countable locally convex topological vector space, where the metric can be derived from the entropy (e.g.
-trough the square root of the Jensen-Shannon divergence).
+While the geometric structures of classical and quantum mechanics are different, their spaces of ensembles have a rather similar structure: they both consist of a bounded convex subset of a metrizable and second-countable locally convex topological vector space, where the metric can be derived from the entropy (e.g.~through the square root of the Jensen-Shannon divergence).
 There are, therefore, three common structures at play: a topology (which physically represents experimental verifiability), a convex structure (which physically represents the ability to make statistical mixtures) and an entropic structure (which physically represents the variability of the elements within each ensemble).
-While the axioms for the topology and the convex structure are readily available, to our knowledge, the correct axioms for the entropic structure still needs to be identified.
+While the axioms for the topology and the convex structure are readily available, to our knowledge, the correct axioms for the entropic structure still need to be identified.
 Moreover, the interplay between the topology, the convex structure and the entropic structure opens additional mathematical questions which have not been investigated.
@@
 As a guide to the reader, let us recall how ensembles are defined in both classical and quantum mechanics.
-For a discrete classical system, the state space $X$ is a countable set with the discrete topology, the space of ensembles consists of all probability measure over the Borel algebra of $X$ and the entropy is given by $S(p_i) = - \sum_i p_i \log p_i$.
+For a discrete classical system, the state space $X$ is a countable set with the discrete topology, the space of ensembles consists of all probability measures over the Borel algebra of $X$ and the entropy is given by $S(p_i) = - \sum_i p_i \log p_i$.
 Note that all logs here are assumed base 2.
@@
 In particular, for a uniform distribution with support $U$, the entropy reduces to $S(\rho_U) = \log \mu(U)$.
-In quantum mechanics, the state space is given by the projective space $P(\mathcal{H})$ of a complex Hilbert space $\mathcal{H}$, the space of ensembles is given by the density operators (i.e.
-positive semi-definite, self-adjoint trace-one operators) and the entropy is given by $S(\rho) = -\tr(\rho \log \rho)$.
+In quantum mechanics, the state space is given by the projective space $P(\mathcal{H})$ of a complex Hilbert space $\mathcal{H}$, the space of ensembles is given by the density operators (i.e.~positive semi-definite, self-adjoint trace-one operators) and the entropy is given by $S(\rho) = -\tr(\rho \log \rho)$.
 
-Note that in all three cases the definition of the entropy is tightly bound to the definition of the space, therefore the biggest problem is understanding how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each cases.
-Fortunately, in our preliminary work we already have found the solution.
-In all cases, the entropy presents an upper bound to its increase under convex combinations: $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$ where $I(p) = -p \log p - (1-p) \log (1-p)$.
-This maximum increase is given precisely when the two ensembles are orthogonal with respect to the inner product of the respective spaces.
-In the classical case, this means that the measures have disjoint support and in the quantum case that the eigenspaces of non-zero eigenvalues are orthogonal.
-In other words, characterizing the interaction of the entropy function with the topology and the convex structure is enough to characterize the entropic structure: the form of $I(p)$, the Shannon entropy formula, and, since in both classical and quantum mechanics ensembles can always be decomposed into orthogonal component, the different entropy formulas can be recovered in each case.
+Note that in all three cases the definition of the entropy is tightly bound to the definition of the space.
+Therefore, the biggest problem is understanding how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each case.
+Fortunately, in our preliminary work we already have found the[a?] solution.
+In all cases, the entropy possesses an upper bound to its increase under convex combination: $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$ where $\bar{p} = 1-p$ and $I(p) = -p \log p - \bar{p}\log \bar{p}$.
+This maximum increase is given precisely when the two ensembles are orthogonal with respect to the inner product of the respective space.
+In the classical case, this means that the measures have disjoint support; in the quantum case, it means that the eigenspaces of non-zero eigenvalues are orthogonal.
+In other words, characterizing the interaction of the entropy function with the topology and the convex structure is enough to characterize the entropic structure.
+That is, the form of $I(p)$---the Shannon entropy formula---can be characterized, and since in both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case.
 
@@
 As we saw in classical mechanics, the symplectic form gives us the Liouville measure which in turn is used to define the entropy.
-Therefore, the entropy of ensemble is defined from the geometry.
+Therefore, the entropy of an ensemble is defined from the geometry of the state space.
 However, note that the entropy for a uniform distribution $\rho_U$ over a set $U$ is given by $S(\rho_U) = \log \mu(U)$.
@@
 Around a point, in fact, we can define Darboux coordinates and define joint probability distributions that are the product of independent marginals along each pair of position and momentum.
-One finds a similar relationships between the entropy of uniform marginal distributions and the integral of the symplectic form over the support of the marginal.
+One finds a similar relationship between the entropy of uniform marginal distributions and the integral of the symplectic form over the support of the marginal.
+[CA: Is this fully equivalent, or should a term other than equivalent be used?]
 
@@
 \end{equation}
-Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, mathematically $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has von Neumann entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi|\psi)}}{2} \right)$ where $I$, as before, is the Shannon entropy for two elements.
+Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has von Neumann entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi|\psi)}}{2} \right)$ where $I$, as before, is the Shannon entropy for two elements.
 As before, the entropy is defined in terms of the inner product.
@@
 
-These insights from our preliminary work shows that the entropic structure plays a crucial role in all physical theories and merits to be studied.
-Moreover, while typically information geometry is cast in finite dimensional spaces, there is nothing preventing its generalization to infinite dimensions.
+Given that \emph{all} physical theories must admit a notion of ensembles that can be statistically mixed via convex combination, these insights from our preliminary work show that the entropic structure plays a crucial role in all physical theories and merits further study.
+Moreover, while typically information geometry is cast in finite-dimensional spaces, there is nothing preventing its generalization to infinite dimensions.
 If the entropic structure, in fact, imposes a distance function whose open balls are convex, then this provides us with a locally convex topology which is exactly what is needed to define calculus in infinite dimensions.
-Moreover, the Lie algebras of both classical (i.e.
-Poisson brackets) and quantum (i.e.
-commutators) have a special relationship with the geometric and entropic structure, which may provide us with a template for generalizes those structure as well.
+Moreover, the Lie algebras of both classical and quantum mechanics, Poisson brackets and commutators, respectively, have a special relationship with the geometric and entropic structures, which may provide us with a template to generalize those structures as well.
 
-Given these very promising preliminary findings, the goal is to find a minimal set of axioms to characterize the entropic structure.
-Minimal in the sense that, in the spirit of analogous work on the foundations of set theory or reverse mathematics, the axioms can be shown to be satisfied by the different theories (which now become model of the larger theory) but also allow to rederive the entropy formula for each specific theory.
+Given these very promising preliminary findings, the goal of the proposed research is to find a minimal set of axioms to characterize the entropic structure of physical theories.
+Minimal in the sense that, in the spirit of analogous work on the foundations of set theory or reverse mathematics, the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory.
 Let us, then, go through the current state of our preliminary work, and show the type of questions that the overall approach touches.
@@
 The central focus is the intuitive notion of statistical ensemble (i.e.~the collection of all outputs of a preparation procedure), as this seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
-In terms of the math, as we have seen, the spaces of statistical ensembles of classical and quantum mechanics can be made to look very similar.
+In terms of the mathematics, as we have seen, the spaces of statistical ensembles of classical and quantum mechanics can be made to look very similar.
 From the experimental side, statistical ensembles are what we produce and measure experimentally.
@@
 Therefore, it makes sense to require that any physical theory must \emph{at least} posit a space of ensembles.
-This provides us with a strong conceptual starting point and it is properties of ensembles that needs to be specified through axioms.
+This provides us with a strong conceptual starting point, and properties of ensembles are what need to be specified through axioms.
 
 The first axiom establishes that an ensemble space $\Ens$ is a $\mathsf{T}_0$ second-countable topological space.
-This is based on our previous work, that established a link between the experimental verifiability of the theory and its topology.
+This is based on our previous work~\cite{aop-book,aop-topExpDisting}, which established a link between the experimental verifiability of the theory and its topology.
 The open sets of the topology correspond to experimentally verifiable statements (i.e.~statements associated with an experimental test that terminates successfully in finite time if and only if the statement is true) while Borel sets correspond to statements associated with an experimental test without termination guarantees~\cite{aop-book,aop-topExpDisting}.
@@
 
-The second axiom imposes a convex structure which represents the ability to perform statistical mixtures.
+The second axiom imposes a convex structure, which represents the ability to perform statistical mixtures.
 Given two ensembles $\ens[a], \ens[b] \in \Ens$, we can always choose a probability $p \in [0,1]$ and select $\ens[a]$ $p$ percent of the times and $\ens[b]$ $\bar{p} = 1-p$ percent of the times.
 This gives the ensemble space a convex structure as $p \ens[a] + \bar{p} \ens[b] \in \Ens$ for all $\ens[a], \ens[b] \in \Ens$ and $p \in [0,1]$.
-This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs) [add refs].
+This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs); see e.g.~Ref.~\cite{plavala2023}.
 The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases).
@@
 Surprisingly, topological convex spaces do not seem to have been studied, and this structure alone presents interesting sets of questions.
-For example, it is established that a convex space embeds into a vector space if and only if it is cancellative.
-That is, if $p \ens[a] + \bar{p} \ens = p \ens[b] + \bar{p} \ens$ for some $p \in (0,1)$ implies $\ens[a] = \ens[b]$.
+For example, it is established that a convex space embeds into a vector space if and only if it is cancellative, that is, if $p \ens[a] + \bar{p} \ens = p \ens[b] + \bar{p} \ens$ for some $p \in (0,1)$ implies $\ens[a] = \ens[b]$.
 The question now is what are necessary and sufficient conditions for a topological convex space to embed continuously into a topological vector space?
-Can we have, for example, an Hausdorff second countable topological convex space that is not a continuous embedding of a topological vector space?
+Can we have, for example, a Hausdorff second-countable topological convex space that is not a continuous embedding of a topological vector space?
 Topological vector spaces present a number of nice results which may or may not hold in topological convex spaces.
-For example, any $\mathsf{T}_1$ second countable topological vector spaces is metrizable.
+For example, any $\mathsf{T}_1$ second-countable topological vector space is metrizable.
 Does that same result hold for topological convex spaces?
@@
 Additionally, there is another type of closure that may be more interesting physically.
-We can define the $\sigma$-hull of a set $A$ the set of possible infinite convex combinations, which can also be shown to be a convex set.
+We can define the $\sigma$-hull of a set $A$ as the set of possible infinite convex combinations, which can also be shown to be a convex set.
 This will be a superset of the hull of $A$, because it will include all finite convex combinations, and it will be a subset of the closed hull of $A$, as all infinite convex combinations are topological limits.
-However, it can be shown that, in some case, the hull, the $\sigma$-hull and the closed hull are all different.
+However, [CA: Remove However?] it can be shown that, in some cases, the hull, the $\sigma$-hull and the closed hull are all different.
 From the physics side, the $\sigma$-hull may be a more physically meaningful closure as it represents the true limit process of statistical mixing, the operation the axiom is characterizing.
@@
 The third axiom, central to the proposed work, imposes an entropy function that quantifies the variability of elements within a statistical ensemble.
-Though we imagine ensembles to be a collection of instances, these instances are never formalized as they need to be left unspecified to leave the theory general.
+Although we imagine an ensemble to be a collection of instances, we do not formalize these instances, as they need to be left unspecified in a general theory.
 However, the variability of the elements within an ensemble is physically significant as it will influence, for example, the variability of statistical measurements.
@@
 
-In our preliminary work, for example, the axiom equips each ensemble space with an entropy function $S : \Ens \to \mathbb{R}$ currently has the following properties:
+In our preliminary work, for example, the axiom equips each ensemble space with an entropy function $S : \Ens \to \mathbb{R}$ that currently has the following properties:
 \begin{itemize}
@@
 The goal is to have a full justification for each property to make sure that we are not leaving out any possibly interesting physical theory.
-Similarly to one does in foundations of mathematics, we typically construct different models of the various formulations of the axioms to check that each property is truly independent from the other, and that is physically required.
+Similarly to what one does in the foundations of mathematics, we typically construct different models of the various formulations of the axioms to check that each property is truly independent from the other, and that it is physically required.
 
-Though we do not believe the axiom lists all necessary properties of the entropic structure, it does list enough of them to already prove interesting results.
+Though we do not believe the current version of the axiom lists all necessary properties of the entropic structure, it does list enough of them to already prove interesting results.
 For example, we can show that $I(p)$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy~\cite{Shannon} $I(p)= - p \log p - \bar{p} \log \bar{p}$.
@@
 Similarly, we will be able to show that, from this structure alone, the geometric structure of both classical and quantum mechanics can be recovered.
+[CA: Above says that we can't recover the symplectic structure from the entropy.]
 
@@
 It is an open question as to whether the ensemble must be a bounded set.
-We can also show that some models of GPTs (i.e.
-Generalized Probabilistic Theories) are inconsistent with an entropic structure and therefore cannot represent a physically meaningful space of ensembles.
-In other words, the entropic structure constrain the type of spaces possible in surprising and interesting ways.
+We can also show that some models of Generalized Probabilistic Theories [add specific refs] are inconsistent with an entropic structure and therefore cannot represent a physically meaningful space of ensembles.
+In other words, the entropic structure constrains the types of spaces possible for physical theories in surprising and interesting ways.
 
@@
 \end{itemize}
-It is almost a distance function, as it does not satisfy the triangle inequality, but it recovers the Jensen-Shannon divergence (JSD) in both classical [add ref] and quantum [add ref] mechanics.
-While the JSD is the square of a distance function in both classical and quantum information theory, our current axioms do not guarantee $MS$ to have this property in general.
+It is almost a distance function, as it does not satisfy the triangle inequality, but it recovers the Jensen-Shannon divergence (JSD)~\cite{Jensen-ShannonOriginal} in both classical and quantum mechanics.
+While the JSD is the square of a distance function in both classical and quantum information theory (see e.g.~\cite{JSDProperties}), our current axioms do not guarantee $MS$ to have this property in general.
 An interesting question, then, is whether enough physical constraints exist for that to be the case.
@@
 
-A related line of questions relates the entropic structure to elements of differential geometry and calculus of variation.
-We are aware that the current axioms do not guarantee the entropic open balls to be convex, even though they are in both the classical and quantum case.
-We suspect that to be the case in general, even though we haven't yet found a physically satisfying premise.
-If the convexity of the open ball is established, then the topology generated by entropic open balls is locally convex, which give us enough structure to define calculus in infinite dimensional spaces.
-That is, the entropic structure would be responsible for the ability to define derivatives.
+A related line of questions relates the entropic structure of a physical theory to elements of differential geometry and the calculus of variations.
+We are aware that the current axioms do not guarantee the entropic open balls to be convex, even though they are in both the classical and quantum cases.
+We suspect the open balls to be convex in general, even though we have not yet found a physically satisfying premise.
+If the convexity of the open balls is established, then the topology generated by entropic open balls is locally convex, which gives us enough structure to define calculus in infinite-dimensional spaces.
+That is, the entropic structure would be responsible for the ability to define derivatives in physical theories.
+[CA: This seems like it would be very powerful.
+Emphasize more somehow?]
 
-In the finite dimensional case, and if the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
+In the finite-dimensional case, and if the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
 Let $V_{\ens[a]}$ be the embedding vector space with $\ens[a] \in \Ens$ as origin.
@@
 Since all Hessians are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is a Hessian manifold.
-In classical and quantum mechanics it recovers the Fisher information metric.
+In classical and quantum mechanics it recovers the Fisher information metric [CA: Add a ref, maybe to an information theory textbook, since having trouble sorting out in which of Fisher's works this metric is introduced].
 The immediate goal, then, is to see how much of information geometry can be generalized to this case.
-If a locally convex topology can be established, the further goal is to extend this to the infinite dimensional case.
+If a locally convex topology can be established, the further goal is to extend this to the infinite-dimensional case.
 
@@
 For classical mechanics, the situation is a bit more complex as we need to recover first the Liouville measure.
-This can be done by noting two things: for a uniform distribution with support $U \subseteq X$, $\mu(U) = 2^{S(\rho_U)}$ and the uniform distribution over $U$ has the highest entropy of all distributions with support $U$.
-Given a set $A \subseteq \Ens$ of ensemble, we can then define the state capacity as supremum of the exponential of the hull of $A$, which physically means the highest entropy achievable with a statistical mixture of elements of $A$.
+This can be done by noting two things: for a uniform distribution with support $U \subseteq X$, $\mu(U) = 2^{S(\rho_U)}$, and the uniform distribution over $U$ has the highest entropy of all distributions with support $U$.
+Given a set $A \subseteq \Ens$ of ensembles, we can then define the state capacity as the supremum of the exponential of the hull of $A$, which physically means the highest entropy achievable with a statistical mixture of elements of $A$.
 This set function has nice properties: it is non-negative, monotonic, subadditive in general and additive on orthogonal sets of ensembles.
@@
 Interestingly enough, one can follow the same construction in the quantum case, and the measure will return the dimensionality of the subspace.
-Concerning non-additive measure, there are more interesting questions as one can create a non-additive version of probability, which may point to a non-additive version of probability theory.
-We have already looked at this problem, and, unfortunately, the standard techniques that have been developed in that area are not sufficient.
+Concerning non-additive measures, there are more interesting questions as one can create a non-additive version of probability, which may point to a non-additive version of probability theory [CA: This seems unclear.].
+We have already looked at this problem, and, unfortunately, the standard techniques that have been developed in that area are not sufficient to handle the cases of interest to us within physical theories.
 
-A possible fourth axiom may be required to conclude all the requirements for physical theories.
+A possible fourth axiom may be needed to conclude all the requirements for physical theories.
 In both classical and quantum mechanics, there is a Lie algebra that characterizes the relationships between generators of time evolution.
-In classical mechanics, this is captured by the Poisson bracket while in quantum mechanics is captured by the commutator.
-Since the set of the probability measures defined on an odd dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the previous three axioms are not sufficient.
-However, there is an intimate relationship between this Lie algebraic structure and the entropic structure since, in both cases, they describe the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged, which means, as outlined before, they must leave the entropic structure unchanged.
+In classical mechanics, this is captured by the Poisson bracket, while in quantum mechanics it is captured by the commutator.
+Since the set of the probability measures defined on an odd-dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the previous three axioms are not sufficient [CA: I don't see how this follows].
+However, there is an intimate relationship between this Lie-algebraic structure and the entropic structure, since in both cases, the Lie-algebraic structure describes the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged.
+This means, as outlined before, they must leave the entropic structure unchanged.
 That is, it describes the group of deterministic and reversible motions, those that leave the entropy unchanged~\cite{aop-HamConsInfoEnt,aop-HamQuantInfo,aop-book}.
-Part of the investigation for the correct specification of the entropic structure, then, will need to consider its impact on the definition of the Lie algebraic structure.
-For example, in our reverse physics approach to classical mechanics we were able to show that a symplectic structure is necessary and sufficient to define an entropy that is invariant under frame transformations.
-If such requirement could be specified in general, the principle of relativity at equal time could be shown to be equivalent to this structure.
+Part of the investigation to identify the correct specification of the entropic structure, then, will need to consider its impact on the definition of the Lie-algebraic structure.
+For example, in our reverse physics approach to classical mechanics~\cite{aop-book,Carcassi:2022bpm}, we were able to show that a symplectic structure is necessary and sufficient to define an entropy that is invariant under frame transformations.
+If such a  requirement could be specified in general, the principle of relativity at equal time could be shown to be equivalent to this structure.
 
@@
 We want to mathematically formalize the intuition that specifying the geometry in physics is equivalent to specify the entropy.
-To do so, we need to give a precise, physically motivated, axiomatic definition of what an entropic structure is for a general theory.
-Then show that this characterized the geometry of the general space in a way that recovers the geometry of classical and quantum mechanics in the specific case.
-While doing that, we will gain insight on how to handle topological issues properly, and hopefully show that ensemble spaces must embed into a second countable metrizable locally convex vector space, so that notion of differentiability can be extended to the infinite dimensional case.
-We will also keep an eye open to connection to measure theory and Lie algebra, so that more features of physical theories can be generalize in the future.
+To do so, we need to give a precise, physically motivated, axiomatic definition of what an entropic structure is for a general physical theory.
+Then we must show that this characterizes the geometry of the general space in a way that recovers the geometry of classical and quantum mechanics in these specific cases.
+While doing that, we will gain insight on how to handle topological issues properly, and hopefully show that ensemble spaces must embed into a second-countable metrizable locally convex vector space, so that the notion of differentiability can be extended to the infinite-dimensional case.
+We will also keep an eye open to connections to measure theory and Lie algebra, so that more features of physical theories can be generalized in the future.
 
 Notes:
+
+-Cite Hall + Reginatto somewhere?
 
@@
 There is no direct overlap with the activities proposed here, and the end date of the grant is moreover before the start date of the work proposed here.
+Co-PI Uribe does not have any current funding.
 \\
 
-Co-PI Uribe does not have any current funding. \\
-
 \noindent
-\textbf{Relationship to other commitments of PI:} PI Aidala has been continuously involved in high-energy experimental QCD since 2001.
+\textbf{Relationship to other commitments of the PI and Co-PI:} PI Aidala has been continuously involved in high-energy experimental QCD since 2001.
 Her separate line of theoretical research on the mathematical foundations of physics is relatively new, with her first publication in 2018~\cite{aop-phys-blueprint}.
@@
 
-TODO: Need to add 1-2 sentences on other commitments of Alejandro as Co-PI?
+\textcolor{red}{Alejandro: Add a few sentences on other commitments.}
 
@@
 Aidala and Uribe will identify suitable journals for publication and conferences at which to disseminate the results, and oversee the direction of the work.
-Uribe will
+Uribe has served as a mathematical consultant for Aidala and Carcassi's broader Assumptions of Physics project since 2018 and has become more involved since 2024, in particular regarding the efforts to develop a generalized theory of ensemble spaces for all physical theories.
+\textcolor{red}{[Alejandro + Christine: Need to figure out how best to articulate Alejandro's role.]}
 
-Carcassi will serve as the technical lead.
+Carcassi will dedicate 20\% of his time to the project and serve as the technical lead.
 He is (currently part-time) staff in the U-M Department of Physics, a Research Area Specialist Senior, with his salary fully supported through grants.
-He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until the graduate student is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.
+\textcolor{red}{[With Alejandro a Co-PI this year, may need to reframe this somewhat.]} He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until the graduate student is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.
 With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
 
-Depending on the interests, background, and skills of the Ph.D.~student, there will be different ways for the student to contribute.
+Depending on the interests, background, and skills of the Ph.D.~student, there will be different ways for the student to contribute to the research.
 One possibility is to analyze what constructions and theorems from topological vector spaces can be generalized to topological convex spaces.
@@
 Yet another possibility, since the JSD is the square of a distance function, is to identify sufficient conditions for a pseudo-distance to be the square of a distance, which may clarify what properties the entropic structure imposes.
+Developing skills in particular in information theory and information geometry will lead to a wide range of possible next steps in the student's career after their Ph.D., with applications for example in quantum computing and in machine learning across academia and the private sector.
 
@@
 One of them is aimed at the level of individuals with upper-level undergraduate training in physics and/or mathematics, and the second, more technical channel is aimed at professional researchers, for which he creates videos presenting detailed technical results or describing open technical questions, leveraging the expertise of others to help advance research goals.
-Graduate student Tobias Thrien originally discovered the AoP project through Carcassi's YouTube channel while a Bachelor's student in Germany several years ago.
-He has since completed a Master's in theoretical cosmology in Germany and is currently a second-year Ph.D.~student at U-M, having come to Michigan specifically to work on the AoP project with the PI and Carcassi.
 Regarding the impact of posting open technical problems, our first collaborative paper that resulted from someone reaching out to us after seeing a video, on classical mechanics as the high-entropy limit of quantum mechanics, is currently under review at a journal~\cite{aop-classicallimit}.
-We were previously completely unaware of the researcher and his work at the University of Innsbruck, Austria.
-Carcassi's August 2025 video on ``The need for physical mathematics'' has already received 49k views and led directly to an invitation to speak at the American Mathematical Society conference in Washington, DC in January 2026.
+Aidala and Carcassi were previously completely unaware of the researcher and his work at the University of Innsbruck, Austria.
+Carcassi's August 2025 video on ``The need for physical mathematics'' has already received 49k views and led directly to an invitation to speak in a special session at the Joint Mathematics Meeting in Washington, DC in January 2026.
 Carcassi additionally writes monthly blog-style essays related to the AoP project and advertised through social media platforms such as Facebook and LinkedIn.
 Thus, the results of the proposed research will reach a much wider audience than through the publications and conference talks alone.
+Results from this project will moreover be added to a future edition of Aidala and Carcassi's open-access book \textit{Assumptions of Physics} published online through Michigan Publishing~\cite{aop-book}.
+As part of making results accessible to others as soon as possible and leveraging input and expertise from others, real-time compilation from the public github repository of the latest draft of the book is set up on the AoP website~\cite{aop-website}.
 Further dissemination of the open-research approach may inspire researchers in completely different areas to adopt a similar approach for their own projects, with potential acceleration of research progress in particular within interdisciplinary areas, where not all necessary expertise may be available within a single research community.
@@
 Finding general mathematical structures that any physical theory must have will also help constrain the search for new physical theories.
-\\
+[Can we make this paragraph more math-focused?]\\
 
@@
 The reach and impact of the school will be assessed by the number of registrants and their demographics, as well as the subsequent views of the recorded lectures.
-The impact of unifying disparate branches of physics will need to be assessed on a longer-term timescale than that of the grant, but early indicators could include talk invitations and paper citations.
+The impact of developing a mathematical framework unifying disparate branches of physics will need to be assessed on a longer-term timescale than that of the grant, but early indicators could include talk invitations and paper citations.
 \\
@@
 
-TODO: Add info for Alejandro.
+\textcolor{red}{Alejandro: Add relevant qualifications.}
 
@@
 Related to the AoP project, he has given seminars to audiences of physicists, mathematicians, and philosophers at places including Harvard, Caltech, Oxford (UK) the University of Pavia (Italy) the Munich Center for Mathematical Philosophy (Germany) and the Wolfram Institute.
-He and Aidala gave a joint public talk for the Michigan State University Advanced Studies Gateway, and he will give a U-M ``Saturday Morning Physics'' public talk in November 2025.
+He and Aidala gave a joint public talk on misconceptions in the foundations of physics for the Michigan State University Advanced Studies Gateway, and he will give a U-M Saturday Morning Physics public talk on ``What is classical and what is quantum?'' in November 2025.
 He was invited and will deliver an overview talk on the AoP project at a special session of the Joint Mathematics Meeting in January 2026.
-[Are all special session talks invited?] His training in computer science and engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from information theory, theoretical computer science, systems theory, signal processing, and operations research.
+\textcolor{red}{[Alejandro: Are all special session talks invited?]} His training in computer science and engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from information theory, theoretical computer science, systems theory, signal processing, and operations research.
 His position as U-M Physics Department staff without teaching responsibilities provides him with time and flexibility to dedicate to research related to the AoP project such as the work proposed here.
-[Some material repetitive--edit.]\\
+See also his biographical sketch included as an Other Supplementary Document.\\
 
 Together, Aidala, Uribe, and Carcassi are uniquely positioned to advance the mathematical foundations of physics, with Aidala's and Uribe's traditional training in physics and mathematics, respectively, complementing Carcassi's formal training in computer science and engineering.
-Carcassi's university staff position has given him the time and flexibility to audit multiple graduate Mathematics and Physics courses at U-M, after having previously audited multiple graduate Physics courses at Stony Brook University.
+Carcassi's university staff position has allowed him to audit multiple graduate Mathematics and Physics courses at U-M, after having previously audited multiple graduate Physics courses at Stony Brook University.
 Uribe has served as a mathematical consultant to Aidala and Carcassi for the broader Assumptions of Physics project since 2018.
@@
 In 2020 Aidala received the U-M John Dewey Award for long-term commitment to the education of undergraduate students.
-Between Aidala and Carcassi, they have mentored more than 50 undergraduate students in research since coming to U-M in 2012, with Carcassi serving as the primary mentor for three undergraduate senior theses related to the AoP project.
-[More on Carcassi's mentoring of undergrads?]\\
+[Split out Gabriele's mentees?] Between Aidala and Carcassi, they have mentored more than 50 undergraduate students in research since coming to U-M in 2012, with Carcassi serving as the primary mentor for three undergraduate senior theses related to the AoP project.
+One of Carcassi's former undergraduate research mentees is now pursuing a Ph.D.
+in Mathematics at UT Austin.\\
 
-TODO: Add info for Alejandro.
+\textcolor{red}{Alejandro: Add info on mentoring.}
 
-Uribe has also served on the dissertation committee of four of Aidala's students thus far.
+Uribe has additionally served on the dissertation committee of four of Aidala's students thus far.
 
-Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with more than 17,600 subscribers currently.
+Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with more than 17,700 subscribers currently.
 He posts a new video monthly, generally targeted at individuals with an advanced undergraduate background in physics and/or mathematics.
@@
 He additionally has a second, more technical YouTube channel (@AssumptionsofPhysicsResearch) aimed at professional researchers, with more than 1600 subscribers.
-He uses this channel as a venue to disseminate research results at a more technical level, including video versions of our preprints or published papers as well as conference presentations.
+He uses this channel as a venue to disseminate research results at a more technical level, including video versions of preprints or published papers as well as conference presentations related to the AoP project.
 He moreover uses the technical channel as an innovative approach to try to ``crowd-source'' technical expertise to advance the research more efficiently.
@@
 We were previously completely unaware of the researcher and his work in Innsbruck, Austria.
-In other forms of reach beyond academia, Carcassi was interviewed in 2024 by Curt Jaimungal for the ``Theories of Everything'' podcast~\cite{Carcassi-ToEInterview}, which has 500k subscribers.
+In other forms of reach beyond academia, Carcassi was interviewed in 2024 by Curt Jaimungal for the ``Theories of Everything'' podcast~\cite{Carcassi-ToEInterview}, which has more than 500k subscribers.
 Carcassi additionally writes monthly blog-style essays related to the AoP project and advertised through social media platforms such as Facebook and LinkedIn.
@@
 He recently performed a slightly updated version of the comedy lecture at the American Physical Society meeting in Chicago in October 2025.
-
-Graduate student Tobias Thrien, for whom partial support is requested, originally discovered the Assumptions of Physics project through Carcassi's YouTube channel while a Bachelor's student in Germany several years ago.
-He has since completed a Master's in theoretical cosmology in Germany and is currently a second-year Ph.D.~student at U-M.
-He attended Aidala and Carcassi's inaugural online summer school on the AoP project in June 2024 and is already highly familiar with the broader project, its goals, and its approaches.
 
@@
 Her QCD research has focused on advancing the multidimensional quark-gluon structure of the proton and on studying the processes by which scattered quarks or gluons form new bound states (hadrons).
-Her group's NSF-funded work includes results published in Nature~\cite{SeaQuest:2021zxb}, Physical Review Letters~\cite{LHCb:2022tbc,LHCb:2019qoc}, Physical Review D Letters~\cite{LHCb:2022rky}, and Physical Review C~\cite{SeaQuest:2022xdu}, among other journals.
+Her group's NSF-funded work includes results published for example in Nature~\cite{SeaQuest:2021zxb}, Physical Review Letters~\cite{LHCb:2022tbc,LHCb:2019qoc}, Physical Review D Letters~\cite{LHCb:2022rky} and Physical Review D~\cite{LHCb:2025mcq}, Physical Review C~\cite{SeaQuest:2022xdu}, Physics Letters B~\cite{LHCb:2025tvf}, and Computing and Software for Big Science~\cite{Aidala:2023dai}, among other journals.
 Regarding the specific scientific advancements of some of the work, Aidala and her collaborators have studied antimatter in the proton and found surprising evidence that there is an excess of down ``flavored'' antiquarks relative to up flavored antiquarks in the proton for antiquarks carrying as much as 45\% of the proton's total momentum~\cite{SeaQuest:2021zxb,SeaQuest:2022xdu}.
-Her group has also led the first-ever measurements of identified hadron species production by high-energy light-mass quarks in proton-proton collisions, multidifferential in multiple kinematic variables simultaneously, showing that existing models poorly predict the relative production rates of some species~\cite{LHCb:2022rky}.
-[TODO: Add dead cone effect] This proposal is the second time she is submitting to the Applied Mathematics program, based on the research direction in the mathematical foundations of physics that she initiated more recently, which has so far produced 14 peer-reviewed papers since 2018 as well as a technical book, as described above in the section on qualifications of the PI and research team.
+Her group also led the first-ever measurements of identified hadron species production by high-energy light-mass quarks in proton-proton collisions, multidifferential in multiple kinematic variables simultaneously, showing that existing models poorly predict the relative production rates of some species~\cite{LHCb:2022rky}, and recently discovered striking evidence of suppressed gluon radiation from beauty (bottom) quarks due to their large mass~\cite{LHCb:2025mcq}.
+In the spirit of making research more open, her group has championed and led the open data efforts of the 1200-scientist LHCb collaboration~\cite{Aidala:2023dai}.
+This proposal is the second time she is submitting to the Applied Mathematics program, based on the research direction in the mathematical foundations of physics that she initiated more recently, which has so far produced 15 peer-reviewed papers since 2018 as well as a technical book, as described above in the section on qualifications of the PI and research team.
 \\
 
-TODO: Prior NSF support for Alejandro?
+\textcolor{red}{Alejandro: Add information on intellectual merit of any prior NSF support.}
 
@@
 
-[TODO: Update and rewrite to avoid DEI language.] In addition, a variety of activities under Aidala's prior NSF grants have contributed directly to training junior scientists and to diversity in the STEM workforce.
-A woman postdoctoral scholar was involved 2019-22.
-Including students with support from on and off of the grant, 8 Ph.D.
-students and 4 undergrads have been involved in the project since July 2020.
+In addition, a variety of activities under Aidala's prior NSF grants have contributed directly to training junior individuals from a wide range of backgrounds for careers in STEM.
+Including individuals with support from on and off of the grant, 3 postdocs, 8 Ph.D.
+students and 6 undergrads have been involved in her NSF-funded projects since July 2020.
 An additional 3 distinct Ph.D.
 students and 7 undergraduates worked on the projects supported by Aidala's CAREER grant 2015-20.
-Of the grad students, 2 have been African-American men, 2 Caucasian men, 2 Latino men, 1 an Arabic man, and 4 women, of which 2 Latina.
-Former African-American grad student Bryan Ramson is now a permanent staff scientist at Fermilab; former Latina grad student Catherine Ayuso is now Head of Special Projects and Research Manager at Entanglement, Inc.
-The others are in postdoctoral positions or still students in the group.
-The undergrads included 1 Latino, 1 man who was openly autistic, and 5 women from Mt.~Holyoke College, among which were international students from the Philippines, Bangladesh, and China.
-Four of the grad/undergrad students openly identify as LGBT.
-All Mt.~Holyoke students have continued in STEM.
-One received an M.S.
-from Stanford in Statistics and now works for Amazon, the other four are now Ph.D.
-students at Harvard, Michigan, Virginia Tech, and Vanderbilt in Applied Physics, Physics, or Electrical Engineering.
+Among her former grad students, Bryan Ramson is now a permanent staff scientist at Fermi National Accelerator Lab, Joseph Osborn a permanent staff scientist at Brookhaven National Lab, Catherine Ayuso Head of Special Projects and Research Manager at Entanglement, Inc., Dylan Manna the owner of Manna Management Consulting, and Ibrahim Chahrour a lecturer at U-M.
+Others are in postdoctoral positions or still students in the group.
 \\
 
-TODO: Prior NSF support for Alejandro?
+\textcolor{red}{Alejandro: Add broader impacts of any prior NSF support.}
 

### 2025-11-03 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Updates with Gabriele

@@
 
-\section{Proposed research program}
+The proposed work starts from the observation that the spaces of statistical ensembles of classical and quantum mechanics have multiple features in common.
+They are both convex subsets of a locally convex topological vector space whose metric is induced by the information entropy, or more specifically, by the Jensen-Shannon divergence.
+This is in contrast to the different structures of the spaces of classical and quantum states, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces, respectively.
+Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can indeed be recovered by the entropy.
+In general, the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
 
-The proposed work starts from the observation that the spaces of statistical ensembles of classical and quantum mechanics have multiple features in common.
-They are both convex subsets of a locally convex topological vector space whose metric is induced by the entropy, or more specifically, by the Jensen-Shannon divergence.
-This is in contrast to the different geometric structures of classical and quantum mechanics, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces, respectively.
-Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can indeed be recovered by the entropy.
-The states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
-
-The goal is to find minimal requirements for an entropic structure that is common to all physical theories in such a way that known results from classical and quantum theories, in areas like mechanics and information theory, can be generalized.
+The goal is to find minimal requirements (i.e.~axioms) for an entropic structure that is common to all physical theories in such a way that known results from classical and quantum theories, in areas like mechanics and information theory, can be generalized.
 In particular, the relationships among the four mathematical structures required by a space of ensembles---the topology, the convex structure, the proposed entropic structure and the Lie algebra---need to be understood.
@@
 
+\section{Proposed activities}
+
 \subsection{Intellectual merit}
 
-The goal of this proposal is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum.
-The immediate aim is to generalize results of both classical and quantum information theory to a common, more abstract theory that can also handle the infinite-dimensional case.
-Longer term, we aim to use the axioms developed here as part of a larger effort to develop a unified base theory for physical theories with a clean mathematical foundation.
-
-While the geometric structures of classical and quantum mechanics are different, their spaces of ensembles have a rather similar structure: they both consist of a bounded convex subset of a metrizable and second-countable locally convex topological vector space, where the metric can be derived from the entropy (e.g.~through the square root of the Jensen-Shannon divergence).
-There are, therefore, three common structures at play: a topology (which physically represents experimental verifiability), a convex structure (which physically represents the ability to make statistical mixtures) and an entropic structure (which physically represents the variability of the elements within each ensemble).
-While the axioms for the topology and the convex structure are readily available, to our knowledge, the correct axioms for the entropic structure still need to be identified.
-Moreover, the interplay between the topology, the convex structure and the entropic structure opens additional mathematical questions which have not been investigated.
-
-As a guide to the reader, let us recall how ensembles are defined in both classical and quantum mechanics.
+As a guide to the reader, let us recall how states and ensembles are defined in both classical and quantum mechanics.
 For a discrete classical system, the state space $X$ is a countable set with the discrete topology, the space of ensembles consists of all probability measures over the Borel algebra of $X$ and the entropy is given by $S(p_i) = - \sum_i p_i \log p_i$.
@@
 
-Note that in all three cases the definition of the entropy is tightly bound to the definition of the space.
+While the structures of the state spaces of classical and quantum mechanics are different, their spaces of ensembles have a rather similar structure: they both consist of a bounded convex subset of a metrizable and second-countable locally convex topological vector space, where the metric can be derived from the entropy through the square root of the Jensen-Shannon divergence $JSD = S\left(\frac{1}{2} \ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2}S(\ens[a]) + \frac{1}{2}S(\ens[b])\right)$~\cite{Jensen-ShannonOriginal}.
+The JSD is, in both cases, the square of a distance function and can be interpreted as how much the average entropy increases under an equal mixture of two ensembles.
+There are, therefore, three common structures at play: a topology (which physically represents experimental verifiability), a convex structure (which physically represents the ability to make statistical mixtures) and an entropic structure (which physically represents the variability of the elements within each ensemble).
+While the axioms for the topology and the convex structure are readily available, to our knowledge, the correct axioms for the entropic structure still need to be identified.
+Moreover, the interplay between the topology, the convex structure and the entropic structure opens additional mathematical questions which have not been investigated.
+
+The goal of this proposed work is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum.
+The immediate aim is to generalize results of both classical and quantum information theory to a common, more abstract theory that can also handle the infinite-dimensional case.
+Longer term, we aim to use the axioms developed here as part of a larger effort to develop a unified base theory for physical theories with a clean mathematical foundation.
+
+Note that in all above three cases the definition of the entropy is tightly bound to the definition of the space.
 Therefore, the biggest problem is understanding how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each case.
-Fortunately, in our preliminary work we already have found the[a?] solution.
+Fortunately, in our preliminary work we already have found the solution.
 In all cases, the entropy possesses an upper bound to its increase under convex combination: $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$ where $\bar{p} = 1-p$ and $I(p) = -p \log p - \bar{p}\log \bar{p}$.
@@
 That is, the entropic structure of the ensemble space is equivalent to the geometric structure of the state space.
+
 As we saw in classical mechanics, the symplectic form gives us the Liouville measure which in turn is used to define the entropy.
@@
 One finds a similar relationship between the entropy of uniform marginal distributions and the integral of the symplectic form over the support of the marginal.
-[CA: Is this fully equivalent, or should a term other than equivalent be used?]
+From these, we believe we can recover the symplectic form.
 
@@
 Given these very promising preliminary findings, the goal of the proposed research is to find a minimal set of axioms to characterize the entropic structure of physical theories.
-Minimal in the sense that, in the spirit of analogous work on the foundations of set theory or reverse mathematics, the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory.
+Minimal in the sense that, in the spirit of analogous work on the foundations of mathematics, like set theory or reverse mathematics (cite revmath book), the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory.
 Let us, then, go through the current state of our preliminary work, and show the type of questions that the overall approach touches.
@@
 This gives the ensemble space a convex structure as $p \ens[a] + \bar{p} \ens[b] \in \Ens$ for all $\ens[a], \ens[b] \in \Ens$ and $p \in [0,1]$.
-This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs); see e.g.~Ref.~\cite{plavala2023}.
-The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases).
+This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs); see e.g.~Ref.~\cite{plavala2023} (cite wilce et al).
+The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases (cite our Hilber and possibly the paper that shows entropy not defined for all distributions)).
 Rather, we let the topology do the work.
@@
 This will be a superset of the hull of $A$, because it will include all finite convex combinations, and it will be a subset of the closed hull of $A$, as all infinite convex combinations are topological limits.
-However, [CA: Remove However?] it can be shown that, in some cases, the hull, the $\sigma$-hull and the closed hull are all different.
+Moreover, it can be shown that, in some cases, the hull, the $\sigma$-hull and the closed hull are all different.
 From the physics side, the $\sigma$-hull may be a more physically meaningful closure as it represents the true limit process of statistical mixing, the operation the axiom is characterizing.
-Therefore, there are already interesting topics to study even without imposing the entropic structure.
+The subtle issue is that convex combinations always return elements ``inside'' the set, while linear combinations can also return the ``edges.'' Therefore, there are already interesting topics to study even without imposing the entropic structure.
 
@@
 For example, if we are performing a statistical mixture between two ensembles, the average variability cannot decrease, which justifies the concavity.
-The goal is to have a full justification for each property to make sure that we are not leaving out any possibly interesting physical theory.
+The goal is to have a full physical justification for each property to make sure that we are not leaving out any possibly interesting physical theory.
 Similarly to what one does in the foundations of mathematics, we typically construct different models of the various formulations of the axioms to check that each property is truly independent from the other, and that it is physically required.
 
-Though we do not believe the current version of the axiom lists all necessary properties of the entropic structure, it does list enough of them to already prove interesting results.
+Though we do not believe the current version of the axiom lists all necessary properties of the entropic structure, as discussed below, it does list enough of them to already prove interesting results.
 For example, we can show that $I(p)$, the maximum entropy increase during mixing, must be, up to a multiplicative constant, the Shannon entropy~\cite{Shannon} $I(p)= - p \log p - \bar{p} \log \bar{p}$.
@@
 Similarly, we will be able to show that, from this structure alone, the geometric structure of both classical and quantum mechanics can be recovered.
-[CA: Above says that we can't recover the symplectic structure from the entropy.]
 
@@
 It is an open question as to whether the ensemble must be a bounded set.
-We can also show that some models of Generalized Probabilistic Theories [add specific refs] are inconsistent with an entropic structure and therefore cannot represent a physically meaningful space of ensembles.
+We believe we can also show that some examples of Generalized Probabilistic Theories are inconsistent with an entropic structure and therefore cannot represent a physically meaningful space of ensembles.
 In other words, the entropic structure constrains the types of spaces possible for physical theories in surprising and interesting ways.
@@
 If the convexity of the open balls is established, then the topology generated by entropic open balls is locally convex, which gives us enough structure to define calculus in infinite-dimensional spaces.
-That is, the entropic structure would be responsible for the ability to define derivatives in physical theories.
-[CA: This seems like it would be very powerful.
-Emphasize more somehow?]
+That is, the entropic structure would in fact be responsible for the ability to define derivatives in physical theories.
 
@@
 On a linear coordinate chart we can define $g_{\ens}( v, w ) = - \frac{\partial^2 S}{\partial \ens ^2}( v, w )$ where $v$ and $w$ are elements of the tangent space at $\ens \in \Ens \subset V_{\ens[a]}$.
-Since all Hessians are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is a Hessian manifold.
+Since all Hessians are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is a Hessian manifold (cite).
 In classical and quantum mechanics it recovers the Fisher information metric [CA: Add a ref, maybe to an information theory textbook, since having trouble sorting out in which of Fisher's works this metric is introduced].
@@
 This set function has nice properties: it is non-negative, monotonic, subadditive in general and additive on orthogonal sets of ensembles.
-Note that if $A_U \subseteq \Ens$ is the subspace of all probability measures with support $U \subseteq X$, the state capacity of $A_U$ recovers the Liouville measure of $I$.
+Note that if $A_U \subseteq \Ens$ is the subspace of all probability measures with support $U \subseteq X$, the state capacity of $A_U$ recovers the Liouville measure of $U$.
 Moreover, $A_U$ is orthogonal to $A_V$ if and only if $U$ and $V$ are disjoint.
@@
 Interestingly enough, one can follow the same construction in the quantum case, and the measure will return the dimensionality of the subspace.
-Concerning non-additive measures, there are more interesting questions as one can create a non-additive version of probability, which may point to a non-additive version of probability theory [CA: This seems unclear.].
+Concerning non-additive measures, there are additional interesting questions as one can create a non-additive version of probability measures, which may point to a non-additive version of probability theory.
 We have already looked at this problem, and, unfortunately, the standard techniques that have been developed in that area are not sufficient to handle the cases of interest to us within physical theories.
@@
 In classical mechanics, this is captured by the Poisson bracket, while in quantum mechanics it is captured by the commutator.
-Since the set of the probability measures defined on an odd-dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the previous three axioms are not sufficient [CA: I don't see how this follows].
+The unsolved difficulty is that these structures are defined on the state space and not on the ensemble space.
+While there are attempts in this direction (cite Reginatto and Ariel), we have found no general solution in the literature.
+Since the set of the probability measures defined on an odd-dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the previous three axioms are not sufficient since no symplectic structure can exist over odd-dimensional manifolds.
 However, there is an intimate relationship between this Lie-algebraic structure and the entropic structure, since in both cases, the Lie-algebraic structure describes the relationship between the generators of transformations that leave the geometric structure (i.e.~symplectic form and inner product respectively) unchanged.
@@
 
-As we have seen, the program is rich and connects to many areas of physics, information theory and mathematics.
+\noindent
+\textbf{Summary of proposed research program:} As we have seen, the program is rich and connects to many areas of physics, information theory and mathematics.
 To not get lost, let us reiterate the primary goals.
-We want to mathematically formalize the intuition that specifying the geometry in physics is equivalent to specify the entropy.
+We want to mathematically formalize the intuition that specifying the geometry in physics is equivalent to specifying the entropy.
 To do so, we need to give a precise, physically motivated, axiomatic definition of what an entropic structure is for a general physical theory.
-Then we must show that this characterizes the geometry of the general space in a way that recovers the geometry of classical and quantum mechanics in these specific cases.
-While doing that, we will gain insight on how to handle topological issues properly, and hopefully show that ensemble spaces must embed into a second-countable metrizable locally convex vector space, so that the notion of differentiability can be extended to the infinite-dimensional case.
+The fundamental grounding concept will be the notion of statistical ensemble, as we can argue that any physical theory must \emph{at least} provide us with a space of statistical ensembles.
+All axioms will be stated as properties of ensemble spaces.
+Then we must show that the entropic structure characterizes the geometry of the general space in a way that recovers the geometry of classical and quantum mechanics in these specific cases.
+While pursuing this work, we will gain insight on how to handle topological issues properly, and hopefully show that ensemble spaces must embed into a second-countable metrizable locally convex vector space, so that the notion of differentiability can be extended to the infinite-dimensional case.
 We will also keep an eye open to connections to measure theory and Lie algebra, so that more features of physical theories can be generalized in the future.
+\\
 
-Notes:
+[Notes:
+Be clearer about how far expect to get in the 3 yrs.
+Papers: 
+1. Recover Shannon entropy as the maximum entropy increase under mixture for all ensemble spaces [can already do]
+2. Show that we can recover the geometric structure of classical and quantum mechanics from the entropy structure alone [can already do]
+3. Show that we can generalize information geometry in finite dimensions [can already do]
+4. Show how the entropic structure constrains the space of physical theories and rules out some of the GPTs [cal already do]
+]
 
--Cite Hall + Reginatto somewhere?
+[Expected significance/benefits: Can have a common axiomatic foundation for all physical theories]
 
--Develop axiomatics for the entropy such that they’re physically motivated.
-Not taking starting points because they’re convenient mathematically, nor because we think one physical theory (e.g.
-quantum mechanics) is the foundational one.
-
--See if this type of framework will help constrain the problem in infinite dimensions (motivation, not goal).
-Need to generalize the Poisson bracket structure in infinite dimensions—relationship between entropy and Poisson brackets used as Lie brackets.
-Well-known open research problem that there’s no general theory for infinite-dimensional analysis/infinite-dimensional Lie groups.
-
-Motivation (but not goal): Eventual framework to generalize classical and quantum information theory
-
-1. To answer fundamental questions.
-
-2. To do classical and quantum info in a more generalized setting
-
-3. Give a structure for infinite-dimensional analysis that’s constrained enough that mathematicians can prove theorems that are useful for physics
-
-To determine which fundamental structures/properties are needed for all physical theories.
-
-Open with: One of the difficulties with mathematical physics is that physicists aren’t good at specifying the mathematical requirements such that mathematicians can work effectively on the problems.
-
-\section{Summary of proposed research program}
-[Add brief recap]
+[Add somewhere: Regarding the relationship of this work to the present state of knowledge in the field, while there exists other work attempting to reconstruct quantum mechanics from physical principles [cite], we are not aware of any literature attempting to relate the geometric and entropic structures of quantum mechanics, much less in a general physical theory.]
+]
 
@@
 \textbf{Assessment of intellectual merit:} \\
+We will know if the proposed research program is successful if [proofs work!].  [Alternatively, if we conclude that these axioms are insufficient.]
 
@@
 \textcolor{red}{Alejandro: Add a few sentences on other commitments.}
-
-\section{Personnel}
-The proposed research will be performed by PI Christine Aidala (U-M Physics), Co-PI Alejandro Uribe (U-M Mathematics), staff researcher Gabriele Carcassi (U-M Physics), and a Ph.D.~student to be recruited from the Applied and Interdisciplinary Mathematics (AIM), Physics, or Mathematics Ph.D.~program.
-See also the section on ``Qualifications of the Research Team'' below.
-\\
-
-Aidala as PI will be responsible for coordinating all of the work.
-She and/or Uribe will serve as the formal academic (co-)advisor for the Ph.D.~student.
-Students in the AIM program are required to have a primary advisor in Mathematics and a co-advisor in another Department.
-Aidala and Uribe will identify suitable journals for publication and conferences at which to disseminate the results, and oversee the direction of the work.
-Uribe has served as a mathematical consultant for Aidala and Carcassi's broader Assumptions of Physics project since 2018 and has become more involved since 2024, in particular regarding the efforts to develop a generalized theory of ensemble spaces for all physical theories.
-\textcolor{red}{[Alejandro + Christine: Need to figure out how best to articulate Alejandro's role.]}
-
-Carcassi will dedicate 20\% of his time to the project and serve as the technical lead.
-He is (currently part-time) staff in the U-M Department of Physics, a Research Area Specialist Senior, with his salary fully supported through grants.
-\textcolor{red}{[With Alejandro a Co-PI this year, may need to reframe this somewhat.]} He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until the graduate student is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.
-With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
-
-Depending on the interests, background, and skills of the Ph.D.~student, there will be different ways for the student to contribute to the research.
-One possibility is to analyze what constructions and theorems from topological vector spaces can be generalized to topological convex spaces.
-This will help us better understand the relationship between the two, and possibly give insights into what properties the entropic structure may impose to bridge the gap.
-Another possibility is to see what results from classical and quantum information theory can be generalized to the entropic structure alone.
-Yet another possibility, since the JSD is the square of a distance function, is to identify sufficient conditions for a pseudo-distance to be the square of a distance, which may clarify what properties the entropic structure imposes.
-Developing skills in particular in information theory and information geometry will lead to a wide range of possible next steps in the student's career after their Ph.D., with applications for example in quantum computing and in machine learning across academia and the private sector.
-
-%\begin{figure*}[htb]
-%	\centering
-%	\includegraphics[scale=.6]{figures/PersonnelTimelineNSF2022.png}
-%	\caption{Estimated timeline of research activities.}
-%	\label{fig:personnelTimeline}
-%\end{figure*}
 
@@
 
+\section{Personnel}
+The proposed research will be performed by PI Christine Aidala (U-M Physics), Co-PI Alejandro Uribe (U-M Mathematics), staff researcher Gabriele Carcassi (U-M Physics), and a Ph.D.~student to be recruited from the Applied and Interdisciplinary Mathematics (AIM), Physics, or Mathematics Ph.D.~program.
+See also the section on ``Qualifications of the Research Team'' below.
+\\
+
+Aidala as PI will be responsible for coordinating all of the work.
+She and/or Uribe will serve as the formal academic (co-)advisor for the Ph.D.~student.
+Students in the AIM program are required to have a primary advisor in Mathematics and a co-advisor in another Department.
+Aidala and Uribe will identify suitable journals for publication and conferences at which to disseminate the results, and oversee the direction of the work.
+Uribe has served as a mathematical consultant for Aidala and Carcassi's broader Assumptions of Physics project since 2018 and has become more involved since 2024, in particular regarding the efforts to develop a generalized theory of ensemble spaces for all physical theories.
+\textcolor{red}{[Alejandro + Christine: Need to figure out how best to articulate Alejandro's role.]}
+
+Carcassi will dedicate 20\% of his time to the project and serve as the technical lead.
+He is (currently part-time) staff in the U-M Department of Physics, a Research Area Specialist Senior, with his salary fully supported through grants.
+\textcolor{red}{[With Alejandro a Co-PI this year, may need to reframe this somewhat.]} He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until the graduate student is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.
+With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
+
+Depending on the interests, background, and skills of the Ph.D.~student, there will be different ways for the student to contribute to the research.
+One possibility is to analyze what constructions and theorems from topological vector spaces can be generalized to topological convex spaces.
+This will help us better understand the relationship between the two, and possibly give insights into what properties the entropic structure may impose to bridge the gap.
+Another possibility is to see what results from classical and quantum information theory can be generalized to the entropic structure alone.
+Yet another possibility, since the JSD is the square of a distance function, is to identify sufficient conditions for a pseudo-distance to be the square of a distance, which may clarify what properties the entropic structure imposes.
+Developing skills in particular in information theory and information geometry will lead to a wide range of possible next steps in the student's career after their Ph.D., with applications for example in quantum computing and in machine learning across academia and the private sector.
+
+%\begin{figure*}[htb]
+%	\centering
+%	\includegraphics[scale=.6]{figures/PersonnelTimelineNSF2022.png}
+%	\caption{Estimated timeline of research activities.}
+%	\label{fig:personnelTimeline}
+%\end{figure*}
+
 \section{Qualifications of the research team}
@@
 
-Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds, and in 2025 received the U-M Rackham Distinguished Graduate Mentor Award.
+Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award ``for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds,'' and in 2025 received the U-M Rackham Distinguished Graduate Mentor Award.
 She has served as the dissertation chair for 10 graduate students in Physics and Applied Physics who have already completed their doctorates, as co-chair for another in Applied Physics and as co-chair with Uribe for one in Applied and Indisciplinary Mathematics.
@@
 In 2020 Aidala received the U-M John Dewey Award for long-term commitment to the education of undergraduate students.
-[Split out Gabriele's mentees?] Between Aidala and Carcassi, they have mentored more than 50 undergraduate students in research since coming to U-M in 2012, with Carcassi serving as the primary mentor for three undergraduate senior theses related to the AoP project.
+Between Aidala and Carcassi, they have mentored more than 50 undergraduate students in research since coming to U-M in 2012, with Carcassi serving as the primary mentor for three undergraduate senior theses related to the AoP program.
 One of Carcassi's former undergraduate research mentees is now pursuing a Ph.D.
@@
 
-Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with more than 17,700 subscribers currently.
+Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with more than 17,800 subscribers currently.
 He posts a new video monthly, generally targeted at individuals with an advanced undergraduate background in physics and/or mathematics.
-Collectively, his videos on this channel have received over 350k views [TODO: update].
+Collectively, his videos on this channel have received over 625k views.
 He additionally has a second, more technical YouTube channel (@AssumptionsofPhysicsResearch) aimed at professional researchers, with more than 1600 subscribers.
@@
 \subsection{Intellectual merit of prior NSF support}
-
 Aidala has been a successful steward of NSF funds since receiving a CAREER Award in 2015.
-Her prior research awards have been through the Physics Division, Nuclear Physics - Experiment program: \emph{CAREER: Valence and sea quark dynamics at Fermilab}, 2015-21, and \emph{Studying quantum chromodynamics at LHCb}, 2020-present (renewed in 2023).
-She is additionally a Co-PI on an NSF AccelNet award, \emph{AccelNet-Design: Inter-American Network of Networks on Quantum Chromodynamics Challenges}, 2021-present, aimed at further developing collaboration in quantum chromodynamics research between the U.S.~and Latin America.
+Her prior research awards have been through the Physics Division, Nuclear Physics - Experiment program: \emph{CAREER: Valence and sea quark dynamics at Fermilab} (Award Number 1452636, Jul 2015-Jun 2021, \$738,082), and \emph{Studying quantum chromodynamics at LHCb} (Award Number 2012926, Jul 2020-Jun 2023, \$697,969; renewal Award Number 2309922, Jul 2023-Jun 2026, \$736,406).
+She is additionally a Co-PI on an NSF AccelNet award, \emph{AccelNet-Design: Inter-American Network of Networks on Quantum Chromodynamics Challenges} (Award Number 2114669, Sep 2021-Aug 2026, \$38,556 allocated to U-M), aimed at further developing collaboration in quantum chromodynamics research between the U.S.~and Latin America.
 Her QCD research has focused on advancing the multidimensional quark-gluon structure of the proton and on studying the processes by which scattered quarks or gluons form new bound states (hadrons).
-Her group's NSF-funded work includes results published for example in Nature~\cite{SeaQuest:2021zxb}, Physical Review Letters~\cite{LHCb:2022tbc,LHCb:2019qoc}, Physical Review D Letters~\cite{LHCb:2022rky} and Physical Review D~\cite{LHCb:2025mcq}, Physical Review C~\cite{SeaQuest:2022xdu}, Physics Letters B~\cite{LHCb:2025tvf}, and Computing and Software for Big Science~\cite{Aidala:2023dai}, among other journals.
+Her group's NSF-funded work includes results published in Nature~\cite{SeaQuest:2021zxb}, Physical Review Letters~\cite{LHCb:2022tbc,LHCb:2019qoc}, Physical Review D Letters~\cite{LHCb:2022rky}, Physical Review D~\cite{LHCb:2025mcq}, Physical Review C~\cite{SeaQuest:2022xdu}, Physics Letters B~\cite{LHCb:2025tvf}, and Computing and Software for Big Science~\cite{Aidala:2023dai}, among other journals~[cite other results--Ezra's, SeaQuest methods, SeaQuest instrumentation], and it includes one additional paper submitted for review~[cite Cynthia's, and possibly ``two additional papers'' if Kara's submitted].
 Regarding the specific scientific advancements of some of the work, Aidala and her collaborators have studied antimatter in the proton and found surprising evidence that there is an excess of down ``flavored'' antiquarks relative to up flavored antiquarks in the proton for antiquarks carrying as much as 45\% of the proton's total momentum~\cite{SeaQuest:2021zxb,SeaQuest:2022xdu}.
@@
 
-\textcolor{red}{Alejandro: Add information on intellectual merit of any prior NSF support.}
+\textcolor{red}{Alejandro: Add information on intellectual merit of any prior NSF support.
+Technically only grants with an end date within the last five years are required to be listed, which I believe you don't have, but since I see you had several NSF grants between 1989 and ending in 2014, I think it would still be good to somehow discuss your prior record of NSF funding, perhaps in broader terms.}
 

### 2025-11-04 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Add citations and some missing text

@@
 
-While the structures of the state spaces of classical and quantum mechanics are different, their spaces of ensembles have a rather similar structure: they both consist of a bounded convex subset of a metrizable and second-countable locally convex topological vector space, where the metric can be derived from the entropy through the square root of the Jensen-Shannon divergence $JSD = S\left(\frac{1}{2} \ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2}S(\ens[a]) + \frac{1}{2}S(\ens[b])\right)$~\cite{Jensen-ShannonOriginal}.
+While the structures of the state spaces of classical and quantum mechanics are different, their spaces of ensembles have a rather similar structure: they both consist of a bounded convex subset of a metrizable and second-countable locally convex topological vector space, where the metric can be derived from the entropy through the square root of the Jensen-Shannon divergence $JSD = S\left(\frac{1}{2} \ens[a] + \frac{1}{2} \ens[b]\right) - \left(\frac{1}{2}S(\ens[a]) + \frac{1}{2}S(\ens[b])\right)$~\cite{Jensen-ShannonOriginal,HayashiQuantumInfo,majtey2005jensen}.
 The JSD is, in both cases, the square of a distance function and can be interpreted as how much the average entropy increases under an equal mixture of two ensembles.
@@
 In other words, characterizing the interaction of the entropy function with the topology and the convex structure is enough to characterize the entropic structure.
-That is, the form of $I(p)$---the Shannon entropy formula---can be characterized, and since in both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case.
+That is, the form of $I(p)$---the Shannon entropy formula~\cite{Shannon}---can be characterized, and since in both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case.
 
@@
 Given these very promising preliminary findings, the goal of the proposed research is to find a minimal set of axioms to characterize the entropic structure of physical theories.
-Minimal in the sense that, in the spirit of analogous work on the foundations of mathematics, like set theory or reverse mathematics (cite revmath book), the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory.
+Minimal in the sense that, in the spirit of analogous work on the foundations of mathematics, like set theory or reverse mathematics~\cite{RevMathBook}, the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory.
 Let us, then, go through the current state of our preliminary work, and show the type of questions that the overall approach touches.
@@
 This gives the ensemble space a convex structure as $p \ens[a] + \bar{p} \ens[b] \in \Ens$ for all $\ens[a], \ens[b] \in \Ens$ and $p \in [0,1]$.
-This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs); see e.g.~Ref.~\cite{plavala2023} (cite wilce et al).
-The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases (cite our Hilber and possibly the paper that shows entropy not defined for all distributions)).
+This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs); see e.g.~Refs.~\cite{plavala2023,wilce2025generalizedprobabilitytheorynotes,DArianoBook} \textcolor{blue}{Anything else to cite?}.
+The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases~\cite{baccetti2013infinite,aop-unphysHilbert}).
 Rather, we let the topology do the work.
@@
 It is almost a distance function, as it does not satisfy the triangle inequality, but it recovers the Jensen-Shannon divergence (JSD)~\cite{Jensen-ShannonOriginal} in both classical and quantum mechanics.
-While the JSD is the square of a distance function in both classical and quantum information theory (see e.g.~\cite{JSDProperties}), our current axioms do not guarantee $MS$ to have this property in general.
+While the JSD is the square of a distance function in both classical and quantum information theory (see e.g.~\cite{JSDProperties,HayashiQuantumInfo}), our current axioms do not guarantee $MS$ to have this property in general.
 An interesting question, then, is whether enough physical constraints exist for that to be the case.
@@
 A related line of questions relates the entropic structure of a physical theory to elements of differential geometry and the calculus of variations.
-We are aware that the current axioms do not guarantee the entropic open balls to be convex, even though they are in both the classical and quantum cases.
+We are aware that the current axioms do not guarantee the entropic open balls to be convex, even though they are convex in both the classical and quantum cases.
 We suspect the open balls to be convex in general, even though we have not yet found a physically satisfying premise.
@@
 On a linear coordinate chart we can define $g_{\ens}( v, w ) = - \frac{\partial^2 S}{\partial \ens ^2}( v, w )$ where $v$ and $w$ are elements of the tangent space at $\ens \in \Ens \subset V_{\ens[a]}$.
-Since all Hessians are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is a Hessian manifold (cite).
-In classical and quantum mechanics it recovers the Fisher information metric [CA: Add a ref, maybe to an information theory textbook, since having trouble sorting out in which of Fisher's works this metric is introduced].
+Since all Hessians are symmetric, $g_{\ens}$ will be bilinear, symmetric and positive definite: it is a metric tensor and $V_{\ens[a]}$ is a Hessian manifold~\cite{hessianmanifoldsShima1997,hessianmanifoldsliu2025}.
+In classical and quantum mechanics it recovers the Fisher information metric (see e.g.~Refs.~\cite{facchi2010,AmariInfoGeom}).
 The immediate goal, then, is to see how much of information geometry can be generalized to this case.
@@
 The unsolved difficulty is that these structures are defined on the state space and not on the ensemble space.
-While there are attempts in this direction (cite Reginatto and Ariel), we have found no general solution in the literature.
+While there are attempts in this direction~\cite{caticha2008,HallReginattoBook}, we have found no general solution in the literature.
 Since the set of the probability measures defined on an odd-dimensional manifold satisfies the previous axioms, at least in the current form, it may be that the previous three axioms are not sufficient since no symplectic structure can exist over odd-dimensional manifolds.
@@
 \noindent
+\textbf{Expected outcomes and significance:}
+During the three-year period of the proposed research, we expect to achieve and submit manuscripts on at least the following:
+\begin{enumerate}
+	\item Demonstrate that the Shannon entropy can be recovered as the maximum entropy increase under mixture for all ensemble spaces, and that the Shannon, Gibbs and von Neumann entropies can be recovered in the specific cases of discrete classical, continuous classical, and quantum mechanics
+	\item Show that the geometric structures of classical and quantum mechanics can be recovered from the entropic structure alone 
+	\item Demonstrate that information geometry can be generalized to the ensemble space of any physical theory of finite dimensions 
+	\item Show that the entropic structure imposes constraints on the space of possible physical theories and rules out some Generalized Probabilistic Theories
+\end{enumerate}
+Additional possible outcomes during or subsequent to the period of proposed activities depending on the rate of advancement include \textcolor{blue}{[list near-term possible outcomes.
+Then discuss possible longer-term future developments stemming from this work]}
+
+Near-term possible: 
+Find what results from information geometry and information theory could be generalized to ensemble spaces.
+\textcolor{blue}{Naturally lends itself to various self-contained contributions by the grad student, undergrads, or individuals outside the project}
+
+Determine whether the convexity of the mixing entropy can be physically motivated and whether that guarantees a continuous embedding into a locally convex topological vector space.
+
+Generalize information geometry to the infinite-dimensional case---would require a locally convex topology to be able to define variations.
+
+Determine whether the ensemble spaces can be in general proven to be metrizable and characterize the connection between the metric and the entropy.
+
+Generalize classical mechanics and classical field theory.
+
+Find a suitable function space that can be used as a model for the case of classical field theories.
+
+The primary significance of the proposed research is that it will \textcolor{blue}{[has the potential to?]} provide a common axiomatic foundation for \emph{all} physical theories.
+Such an axiomatic foundation common to all physical theories will serve not only to better ground current theories and elucidate their relationships to one another, but also to constrain potential future physical theories, and perhaps point to assumptions or structures in current theories that hold promise for modification in order to develop new theories.
+
+[We don't know where the mathematical structures in physics come from.
+Therefore, when people try to generalize them, they do not know in which direction they can be generalized.
+What we're doing is instead finding axioms from physical requirements---things the physics must at least give you.
+The axiomatics are justified from the physics.
+We're doing the due diligence to check that the axioms can be justified from the physics and they're independent of each other.
+So when we're doing the generalization, the generalization is better constrained by the physics.]
+
+Regarding the relationship of this work to the present state of knowledge in the field, while there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\textcolor{blue}{[cite Hardy, Ariel, Pavia, see 2:36 into talk by Nicolas Medina Sanchez (Vienna) at Perimeter], we are not aware of any literature attempting to relate the geometric and entropic structures of quantum mechanics, much less in a general physical theory.
+One sentence about general approach(es) of those works?
+There are people trying to reconstruct QM from physical premises, but these works are typically on finite dimensions and do not generalize to CM, and definitely don't generalize to classical or quantum field theories.}
+
+\noindent
 \textbf{Summary of proposed research program:} As we have seen, the program is rich and connects to many areas of physics, information theory and mathematics.
-To not get lost, let us reiterate the primary goals.
+Let us reiterate the primary goals.
 We want to mathematically formalize the intuition that specifying the geometry in physics is equivalent to specifying the entropy.
@@
 \\
-
-[Notes:
-Be clearer about how far expect to get in the 3 yrs.
-Papers: 
-1. Recover Shannon entropy as the maximum entropy increase under mixture for all ensemble spaces [can already do]
-2. Show that we can recover the geometric structure of classical and quantum mechanics from the entropy structure alone [can already do]
-3. Show that we can generalize information geometry in finite dimensions [can already do]
-4. Show how the entropic structure constrains the space of physical theories and rules out some of the GPTs [cal already do]
-]
-
-[Expected significance/benefits: Can have a common axiomatic foundation for all physical theories]
-
-[Add somewhere: Regarding the relationship of this work to the present state of knowledge in the field, while there exists other work attempting to reconstruct quantum mechanics from physical principles [cite], we are not aware of any literature attempting to relate the geometric and entropic structures of quantum mechanics, much less in a general physical theory.]
-]
 
@@
 
+\textcolor{blue}{Mention that we expect to include undergrads}
+
 %\begin{figure*}[htb]
@@
 Her QCD research has focused on advancing the multidimensional quark-gluon structure of the proton and on studying the processes by which scattered quarks or gluons form new bound states (hadrons).
-Her group's NSF-funded work includes results published in Nature~\cite{SeaQuest:2021zxb}, Physical Review Letters~\cite{LHCb:2022tbc,LHCb:2019qoc}, Physical Review D Letters~\cite{LHCb:2022rky}, Physical Review D~\cite{LHCb:2025mcq}, Physical Review C~\cite{SeaQuest:2022xdu}, Physics Letters B~\cite{LHCb:2025tvf}, and Computing and Software for Big Science~\cite{Aidala:2023dai}, among other journals~[cite other results--Ezra's, SeaQuest methods, SeaQuest instrumentation], and it includes one additional paper submitted for review~[cite Cynthia's, and possibly ``two additional papers'' if Kara's submitted].
+Her group's NSF-funded work includes results published in Nature~\cite{SeaQuest:2021zxb}, Physical Review Letters~\cite{LHCb:2022tbc,LHCb:2019qoc}, Physical Review D Letters~\cite{LHCb:2022rky}, Physical Review D~\cite{LHCb:2025mcq}, Physical Review C~\cite{SeaQuest:2022xdu}, Physics Letters B~\cite{LHCb:2025tvf,SeaQuest:2024qdw}, and Computing and Software for Big Science~\cite{Aidala:2023dai}, among other journals~\cite{SeaQuest:2023tcr,SeaQuest:2017kjt}, and it includes one additional paper submitted for review~\cite{LHCb:2025rxf}.
 Regarding the specific scientific advancements of some of the work, Aidala and her collaborators have studied antimatter in the proton and found surprising evidence that there is an excess of down ``flavored'' antiquarks relative to up flavored antiquarks in the proton for antiquarks carrying as much as 45\% of the proton's total momentum~\cite{SeaQuest:2021zxb,SeaQuest:2022xdu}.

### 2025-11-04 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Added a few more refs and changed text that needs work to blue

@@
 \end{enumerate}
-Additional possible outcomes during or subsequent to the period of proposed activities depending on the rate of advancement include \textcolor{blue}{[list near-term possible outcomes.
+\textcolor{blue}{Additional possible outcomes during or subsequent to the period of proposed activities depending on the rate of advancement include [list near-term possible outcomes.
 Then discuss possible longer-term future developments stemming from this work]}
 
-Near-term possible: 
+\textcolor{blue}{Near-term possible: 
 Find what results from information geometry and information theory could be generalized to ensemble spaces.
-\textcolor{blue}{Naturally lends itself to various self-contained contributions by the grad student, undergrads, or individuals outside the project}
+Naturally lends itself to various self-contained contributions by the grad student, undergrads, or individuals outside the project}
 
-Determine whether the convexity of the mixing entropy can be physically motivated and whether that guarantees a continuous embedding into a locally convex topological vector space.
+\textcolor{blue}{Determine whether the convexity of the mixing entropy can be physically motivated and whether that guarantees a continuous embedding into a locally convex topological vector space.}
 
-Generalize information geometry to the infinite-dimensional case---would require a locally convex topology to be able to define variations.
+\textcolor{blue}{Generalize information geometry to the infinite-dimensional case---would require a locally convex topology to be able to define variations.}
 
-Determine whether the ensemble spaces can be in general proven to be metrizable and characterize the connection between the metric and the entropy.
+\textcolor{blue}{Determine whether the ensemble spaces can be in general proven to be metrizable and characterize the connection between the metric and the entropy.}
 
-Generalize classical mechanics and classical field theory.
+\textcolor{blue}{Generalize classical mechanics and classical field theory.}
 
-Find a suitable function space that can be used as a model for the case of classical field theories.
+\textcolor{blue}{Find a suitable function space that can be used as a model for the case of classical field theories.}
 
-The primary significance of the proposed research is that it will \textcolor{blue}{[has the potential to?]} provide a common axiomatic foundation for \emph{all} physical theories.
-Such an axiomatic foundation common to all physical theories will serve not only to better ground current theories and elucidate their relationships to one another, but also to constrain potential future physical theories, and perhaps point to assumptions or structures in current theories that hold promise for modification in order to develop new theories.
+The primary significance of the proposed research is that it will \textcolor{blue}{[has the potential to?] provide a common axiomatic foundation for \emph{all} physical theories.
+Such an axiomatic foundation common to all physical theories will serve not only to better ground current theories and elucidate their relationships to one another, but also to constrain potential future physical theories, and perhaps point to assumptions or structures in current theories that hold promise for modification in order to develop new theories.}
 
-[We don't know where the mathematical structures in physics come from.
+\textcolor{blue}{[Rewrite above paragraph along these lines] We don't know where the mathematical structures in physics come from.
 Therefore, when people try to generalize them, they do not know in which direction they can be generalized.
@@
 We're doing the due diligence to check that the axioms can be justified from the physics and they're independent of each other.
-So when we're doing the generalization, the generalization is better constrained by the physics.]
+So when we're doing the generalization, the generalization is better constrained by the physics.}
 
-Regarding the relationship of this work to the present state of knowledge in the field, while there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\textcolor{blue}{[cite Hardy, Ariel, Pavia, see 2:36 into talk by Nicolas Medina Sanchez (Vienna) at Perimeter], we are not aware of any literature attempting to relate the geometric and entropic structures of quantum mechanics, much less in a general physical theory.
+Regarding the relationship of this work to the present state of knowledge in the field, while there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature \textcolor{blue}{attempting to relate the geometric and entropic structures of quantum mechanics, much less in a general physical theory.
 One sentence about general approach(es) of those works?
@@
 \textbf{Assessment of intellectual merit:} \\
-We will know if the proposed research program is successful if [proofs work!].  [Alternatively, if we conclude that these axioms are insufficient.]
+\textcolor{blue}{We will know if the proposed research program is successful if [proofs work!].  [Alternatively, if we conclude that these axioms are insufficient.]}
 

### 2025-11-04 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Changes from while talking to Alejandro

@@
 Her separate line of theoretical research on the mathematical foundations of physics is relatively new, with her first publication in 2018~\cite{aop-phys-blueprint}.
-She has successfully split her efforts between her different research lines for the past seven years, with her work related to the Assumptions of Physics project producing 15 peer-reviewed publications, 1 more submitted paper, and a published technical book since 2018.
+She has successfully split her efforts between her different research lines for the past seven years, with her work related to the Assumptions of Physics program producing 15 peer-reviewed publications, 1 more submitted paper, and a published technical book since 2018.
 She is moreover in the process of reducing the number of personnel she supervises on QCD research in order to have more time for her theoretical work---four of her Ph.D.~students in QCD completed their degrees in the last 18 months, and two more are expected to graduate before the start of the proposed project.
@@
 Finding general mathematical structures that any physical theory must have will also help constrain the search for new physical theories.
-[Can we make this paragraph more math-focused?]\\
+\textcolor{blue}{[Can we make this paragraph more math-focused?]}\\
 
@@
 Aidala and Uribe will identify suitable journals for publication and conferences at which to disseminate the results, and oversee the direction of the work.
-Uribe has served as a mathematical consultant for Aidala and Carcassi's broader Assumptions of Physics project since 2018 and has become more involved since 2024, in particular regarding the efforts to develop a generalized theory of ensemble spaces for all physical theories.
-\textcolor{red}{[Alejandro + Christine: Need to figure out how best to articulate Alejandro's role.]}
+Uribe has served as a mathematical consultant for Aidala and Carcassi's broader Assumptions of Physics program since 2018 and has become more involved since 2024, in particular regarding the efforts to develop a generalized theory of ensemble spaces for all physical theories.
+He is bringing mathematical expertise and will be able to contextualize the results of the project with respect to other areas within mathematical physics.
+He is interested in what mathematical research questions this project may generate and what interesting mathematical structures may emerge.
 
-Carcassi will dedicate 20\% of his time to the project and serve as the technical lead.
+[Aidala, Uribe, and Carcassi, and once identified, the grad student, will generally meet weekly to discuss the project]
+
+Carcassi will dedicate 20\% of his time to the project and has been the individual most immersed in the technical details of the preliminary work leading to this proposal.
 He is (currently part-time) staff in the U-M Department of Physics, a Research Area Specialist Senior, with his salary fully supported through grants.
-\textcolor{red}{[With Alejandro a Co-PI this year, may need to reframe this somewhat.]} He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until the graduate student is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.
-With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
+\textcolor{blue}{He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until the graduate student is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.}  With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
 
@@
 He and Aidala gave a joint public talk on misconceptions in the foundations of physics for the Michigan State University Advanced Studies Gateway, and he will give a U-M Saturday Morning Physics public talk on ``What is classical and what is quantum?'' in November 2025.
-He was invited and will deliver an overview talk on the AoP project at a special session of the Joint Mathematics Meeting in January 2026.
-\textcolor{red}{[Alejandro: Are all special session talks invited?]} His training in computer science and engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from information theory, theoretical computer science, systems theory, signal processing, and operations research.
+He was invited and will deliver an overview talk on the AoP program at a special session of the Joint Mathematics Meeting in January 2026.
+His training in computer science and engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from information theory, theoretical computer science, systems theory, signal processing, and operations research.
 His position as U-M Physics Department staff without teaching responsibilities provides him with time and flexibility to dedicate to research related to the AoP project such as the work proposed here.

### 2025-11-04 Proposals/2025-NSF-EntropicGeometry/ProjectSummary.tex
Updates with Gabriele
Changes from while talking to Alejandro

@@
 \textbf{Overview:}\\
+\textcolor{blue}{[Provide brief overview]}
 
@@
 \textbf{Intellectual Merit:}  \\
+We want to mathematically formalize the intuition that specifying the geometry in physics is equivalent to specifying the entropy.
+To do so, we need to give a precise, physically motivated, axiomatic definition of what an entropic structure is for a general physical theory.
+The fundamental grounding concept will be the notion of statistical ensemble, as we can argue that any physical theory must \emph{at least} provide us with a space of statistical ensembles.
+All axioms will be stated as properties of ensemble spaces.
+Then we must show that the entropic structure characterizes the geometry of the general space in a way that recovers the geometry of classical and quantum mechanics in these specific cases.
+While pursuing this work, we will gain insight on how to handle topological issues properly, and hopefully show that ensemble spaces must embed into a second-countable metrizable locally convex vector space, so that the notion of differentiability can be extended to the infinite-dimensional case.
+We will also keep an eye open to connections to measure theory and Lie algebra, so that more features of physical theories can be generalized in the future.
+\\
 
 \noindent
-\textbf{Broader impacts:}  \\
-
-The proposed activities additionally include significant outreach through social media, primarily through YouTube.
+\textbf{Broader Impacts:} 
+The proposed activities additionally include an annual online summer school and significant outreach through social media, primarily through YouTube.
 Videos will be produced to disseminate the results of the research, with some videos aimed at the level of an advanced undergraduate in mathematics and/or physics, furthering education.

### 2025-11-07 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Further edits after discussion with G.

@@
 \textbf{Expected outcomes and significance:}
-During the three-year period of the proposed research, we expect to achieve and submit manuscripts on at least the following:
-\begin{enumerate}
-	\item Demonstrate that the Shannon entropy can be recovered as the maximum entropy increase under mixture for all ensemble spaces, and that the Shannon, Gibbs and von Neumann entropies can be recovered in the specific cases of discrete classical, continuous classical, and quantum mechanics
-	\item Show that the geometric structures of classical and quantum mechanics can be recovered from the entropic structure alone 
-	\item Demonstrate that information geometry can be generalized to the ensemble space of any physical theory of finite dimensions 
-	\item Show that the entropic structure imposes constraints on the space of possible physical theories and rules out some Generalized Probabilistic Theories
-\end{enumerate}
-\textcolor{blue}{Additional possible outcomes during or subsequent to the period of proposed activities depending on the rate of advancement include [list near-term possible outcomes.
-Then discuss possible longer-term future developments stemming from this work]}
+During the three-year period of the proposed research, we expect to achieve \textcolor{blue}{and submit manuscripts on} at least the following:
+\begin{itemize}
+	\item Demonstrate that the Shannon entropy can be recovered as the maximum entropy increase under mixture for all ensemble spaces, and that the Shannon, Gibbs and von Neumann entropies can be recovered in the specific cases of discrete classical, continuous classical, and quantum mechanics.
+	\item Show that the geometric structures of classical and quantum mechanics can be recovered from the entropic structure alone.
+	\item Demonstrate that information geometry can be generalized to the ensemble space of any physical theory of finite dimensions. 
+	\item Show that the entropic structure imposes constraints on the space of possible physical theories and rules out some Generalized Probabilistic Theories.
+\end{itemize}
+The above expected results should lead naturally to a range of additional possible directions and questions that could be addressed during or subsequent to the grant period, depending on the rate of advancement.
+For example:
+\begin{itemize}
+	\item Identify what results from information geometry and information theory could be generalized to ensemble spaces; this would lend itself to various self-contained contributions by the grad student, undergrads, or individuals outside the project.
+	\item Determine whether the convexity of the mixing entropy can be physically motivated and whether that guarantees a continuous embedding into a locally convex topological vector space.
+	\item Generalize information geometry to the infinite-dimensional case; this would require a locally convex topology to be able to define variations.
+	\item Determine whether the ensemble spaces can be in general proven to be metrizable and characterize the connection between the metric and the entropy.
+	\item \textcolor{blue}{Generalize classical mechanics and classical field theory.}
+	\item Find a suitable function space that can be used as a model for the case of classical field theories.
+\end{itemize}
 
-\textcolor{blue}{Near-term possible: 
-Find what results from information geometry and information theory could be generalized to ensemble spaces.
-Naturally lends itself to various self-contained contributions by the grad student, undergrads, or individuals outside the project}
+\textcolor{blue}{The primary significance of the proposed research lies in its approach as well as its specific potential to provide a common axiomatic foundation for \emph{all} physical theories.
+Currently within physics, we typically do not know from where the various mathematical structures emerge, making it unclear in which directions they can be generalized in order to develop new theories that are physically meaningful.
+Our approach instead seeks to \emph{find axioms from physical requirements}, i.e.~the axiomatics are justified from the physics.
+Such an axiomatic foundation common to all physical theories would serve not only to better ground current theories and elucidate their relationships to one another, but also to constrain potential future physical theories}
 
-\textcolor{blue}{Determine whether the convexity of the mixing entropy can be physically motivated and whether that guarantees a continuous embedding into a locally convex topological vector space.}
+%We're checking that the axioms are independent of each other.
 
-\textcolor{blue}{Generalize information geometry to the infinite-dimensional case---would require a locally convex topology to be able to define variations.}
-
-\textcolor{blue}{Determine whether the ensemble spaces can be in general proven to be metrizable and characterize the connection between the metric and the entropy.}
-
-\textcolor{blue}{Generalize classical mechanics and classical field theory.}
-
-\textcolor{blue}{Find a suitable function space that can be used as a model for the case of classical field theories.}
-
-The primary significance of the proposed research is that it will \textcolor{blue}{[has the potential to?] provide a common axiomatic foundation for \emph{all} physical theories.
-Such an axiomatic foundation common to all physical theories will serve not only to better ground current theories and elucidate their relationships to one another, but also to constrain potential future physical theories, and perhaps point to assumptions or structures in current theories that hold promise for modification in order to develop new theories.}
-
-\textcolor{blue}{[Rewrite above paragraph along these lines] We don't know where the mathematical structures in physics come from.
-Therefore, when people try to generalize them, they do not know in which direction they can be generalized.
-What we're doing is instead finding axioms from physical requirements---things the physics must at least give you.
-The axiomatics are justified from the physics.
-We're doing the due diligence to check that the axioms can be justified from the physics and they're independent of each other.
-So when we're doing the generalization, the generalization is better constrained by the physics.}
-
-Regarding the relationship of this work to the present state of knowledge in the field, while there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature \textcolor{blue}{attempting to relate the geometric and entropic structures of quantum mechanics, much less in a general physical theory.
-One sentence about general approach(es) of those works?
-There are people trying to reconstruct QM from physical premises, but these works are typically on finite dimensions and do not generalize to CM, and definitely don't generalize to classical or quantum field theories.}
+Regarding the relationship of this work to the present state of knowledge in the field, while there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature \textcolor{blue}{attempting to relate the geometric and entropic structures of quantum mechanics, much less those of a general physical theory.
+Works trying to reconstruct quantum mechancis from physical premises are typically in finite dimensions, do not generalize to classical mechanics, and do not provide any path to generalize to classical or quantum field theories [add refs or recite some?].}
 
@@
 \textbf{Assessment of intellectual merit:} \\
-\textcolor{blue}{We will know if the proposed research program is successful if [proofs work!].  [Alternatively, if we conclude that these axioms are insufficient.]}
+\textcolor{blue}{The proposed research program will be evaluated by the progress made in achieving the stated goals and outcomes.  }
 
@@
 
-[Aidala, Uribe, and Carcassi, and once identified, the grad student, will generally meet weekly to discuss the project]
-
 Carcassi will dedicate 20\% of his time to the project and has been the individual most immersed in the technical details of the preliminary work leading to this proposal.
 He is (currently part-time) staff in the U-M Department of Physics, a Research Area Specialist Senior, with his salary fully supported through grants.
-\textcolor{blue}{He will identify the most promising technical directions, perform a significant fraction of the calculations, serve as the lead writer for publications until the graduate student is at a level to do so, and connect the results of the project to broader results so that the impact of the work is maximized.}  With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
+He, Uribe, and Aidala will together identify the most promising technical directions, he will take the lead in performing a significant fraction of the work, and he will connect the results of the project to broader results so that the impact of the research is maximized.
+With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
+Aidala, Uribe, and Carcassi, and once identified, the Ph.D.~student, will generally meet weekly to discuss the project.
 
@@
 Developing skills in particular in information theory and information geometry will lead to a wide range of possible next steps in the student's career after their Ph.D., with applications for example in quantum computing and in machine learning across academia and the private sector.
-
-\textcolor{blue}{Mention that we expect to include undergrads}
+We additionally anticipate involving one or more undergraduates in the project, identifying self-contained pieces of the project to which they could contribute.
+\textcolor{blue}{For example, starting from a few specific results from information geometry or information theory, they could work on trying generalize these results to ensemble spaces.}
 

### 2025-11-08 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Various further updates

@@
 Physics is currently a patchwork of different theories with no common underlying foundation.
-As part of our larger project, Assumptions of Physics, we aim to develop an axiomatic basis for states and processes that can be equally applied to different theories, such as classical and quantum mechanics.
+As part of our larger Assumptions of Physics program, we aim to develop an axiomatic basis for states and processes that can be equally applied to different theories, such as classical and quantum mechanics.
 Different theories are then recovered as specializations of the more general structure.
@@
 We can, in fact, produce a model of that case and see exactly why the entropic structure fails.
-We can also show ensemble spaces must be ``directionally bounded,'' meaning its intersection with any affine line is a finite interval.
-It is an open question as to whether the ensemble must be a bounded set.
+We can also show that an ensemble space must be ``directionally bounded,'' meaning its intersection with any affine line is a finite interval.
+It is an open question as to whether the ensemble space must be a bounded set.
 We believe we can also show that some examples of Generalized Probabilistic Theories are inconsistent with an entropic structure and therefore cannot represent a physically meaningful space of ensembles.
@@
 	\item Determine whether the ensemble spaces can be in general proven to be metrizable and characterize the connection between the metric and the entropy.
-	\item \textcolor{blue}{Generalize classical mechanics and classical field theory.}
 	\item Find a suitable function space that can be used as a model for the case of classical field theories.
+	\item Generalize classical Hamiltonian mechanics and the principle of stationary action to a generic ensemble space.
 \end{itemize}
 
-\textcolor{blue}{The primary significance of the proposed research lies in its approach as well as its specific potential to provide a common axiomatic foundation for \emph{all} physical theories.
+\textcolor{blue}{The immediate significance of the proposed research stems from the realization that, though classical information theory and quantum information theory are presented as two separate subjects in textbooks, there are a lot of commonalities that make them an interesting target for generalization.
+The broader significance lies in its approach as well as its specific potential to provide a common axiomatic foundation for \emph{all} physical theories.
 Currently within physics, we typically do not know from where the various mathematical structures emerge, making it unclear in which directions they can be generalized in order to develop new theories that are physically meaningful.
 Our approach instead seeks to \emph{find axioms from physical requirements}, i.e.~the axiomatics are justified from the physics.
-Such an axiomatic foundation common to all physical theories would serve not only to better ground current theories and elucidate their relationships to one another, but also to constrain potential future physical theories}
+Such an axiomatic foundation common to all physical theories would serve not only to better ground current theories and elucidate their relationships, but also to constrain potential future physical theories.}
 
@@
 
-Regarding the relationship of this work to the present state of knowledge in the field, while there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature \textcolor{blue}{attempting to relate the geometric and entropic structures of quantum mechanics, much less those of a general physical theory.
-Works trying to reconstruct quantum mechancis from physical premises are typically in finite dimensions, do not generalize to classical mechanics, and do not provide any path to generalize to classical or quantum field theories [add refs or recite some?].}
+While there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature attempting to relate the geometric and entropic structures of either classical or quantum mechanics, much less those of a general physical theory.
+Works trying to reconstruct quantum mechancis from physical premises are typically in finite dimensions, do not generalize to classical mechanics, and do not provide a mathematically rigorous path to generalize to classical or quantum field theories.
 
@@
 We will also keep an eye open to connections to measure theory and Lie algebra, so that more features of physical theories can be generalized in the future.
-\\
 
 \noindent
-\textbf{Assessment of intellectual merit:} \\
-\textcolor{blue}{The proposed research program will be evaluated by the progress made in achieving the stated goals and outcomes.  }
+\textbf{Assessment of intellectual merit:} The proposed research program will be evaluated by the progress made in achieving the stated goals and outcomes and in disseminating the results to technical and academic audiences.
 
@@
 Her grant from the John Templeton Foundation, titled ``Why classical probability and classical information theory are incompatible with quantum mechanics and quantum contextuality,'' supports the development of novel non-additive measures (set functions) suitable for quantifying the number of states in quantum mechanics~\cite{aop-nonaddmeas}.
-There is no direct overlap with the activities proposed here, and the end date of the grant is moreover before the start date of the work proposed here.
+There is no direct overlap with the activities proposed here.
 Co-PI Uribe does not have any current funding.
-\\
 
 \noindent
-\textbf{Relationship to other commitments of the PI and Co-PI:} PI Aidala has been continuously involved in high-energy experimental QCD since 2001.
+\textbf{Relationship to other commitments and activities of the PI and Co-PI:} PI Aidala has been continuously involved in high-energy experimental QCD since 2001.
 Her separate line of theoretical research on the mathematical foundations of physics is relatively new, with her first publication in 2018~\cite{aop-phys-blueprint}.
@@
 
-\textcolor{red}{Alejandro: Add a few sentences on other commitments.}
+Co-PI Uribe has an active research program studying the semi-classical limit in various geometric contexts, currently with a focus on phase spaces with a K\"{a}hler structure.
+This limit relates the mathematics of quantum and classical mechanics, as Planck's constant tends to zero.
+There is a similarity between Uribe's research program and the research proposed here, as the latter unifies the formalisms of classical and quantum mechanics.
+Uribe is very interested in exploring this similarity, as it may give rise to new perspectives and suggest new theorems.
 
 \subsection{Broader impacts}
-In addition to disseminating the goals and results of the project through traditional means such as peer-reviewed publications and presentations at workshops and conferences, the goals and results of the project will be disseminated to wider audiences by leveraging existing social media and other outreach infrastructure that has already been developed for the broader Assumptions of Physics project, in particular, Carcassi's two YouTube channels, as described below in the section on qualifications of the research team.
+In addition to disseminating the goals and results of the project through traditional means such as peer-reviewed publications and presentations at workshops and conferences, the goals and results of the project will be disseminated to wider audiences by leveraging existing social media and other outreach infrastructure that has already been developed for the broader Assumptions of Physics program, in particular, Carcassi's two YouTube channels, as described below in the section on qualifications of the research team.
 One of them is aimed at the level of individuals with upper-level undergraduate training in physics and/or mathematics, and the second, more technical channel is aimed at professional researchers, for which he creates videos presenting detailed technical results or describing open technical questions, leveraging the expertise of others to help advance research goals.
@@
 Carcassi's August 2025 video on ``The need for physical mathematics'' has already received 49k views and led directly to an invitation to speak in a special session at the Joint Mathematics Meeting in Washington, DC in January 2026.
-Carcassi additionally writes monthly blog-style essays related to the AoP project and advertised through social media platforms such as Facebook and LinkedIn.
+Carcassi additionally writes monthly blog-style essays related to the AoP program and advertised through social media platforms such as Facebook and LinkedIn.
 Thus, the results of the proposed research will reach a much wider audience than through the publications and conference talks alone.
@@
 
-Carcassi and Aidala will host an annual online school that includes an introduction to generalized ensemble spaces and the mathematical relationships between geometric and entropic structures, showcasing various results from the proposed work in a didactic manner.
+As part of this project, Carcassi and Aidala will host an annual online school that includes an introduction to generalized ensemble spaces and the mathematical relationships between geometric and entropic structures, showcasing various results from the proposed work in a didactic manner.
 The school will consist of six hour-long lectures, two per day over three days, each followed by 30 minutes of questions and discussion, from 10 am - 1 pm U.S.
@@
 
-Beyond the immediate research goals, the proposed work will serve to unify what are currently disparate branches of physics---classical and quantum, and aspects of statistical mechanics.
-Developing a common framework will pave the way for eventual new approaches to physics education, with less compartmentalized teaching across the core advanced undergraduate courses.
+Beyond the immediate research goals, the proposed work will serve to unify what are currently disparate branches of physics---classical and quantum, and aspects of statistical mechanics---while making them more mathematically rigorous.
+Developing a common mathematical framework will pave the way for eventual new approaches to physics education, with less compartmentalized teaching across the core advanced undergraduate courses.
 Finding general mathematical structures that any physical theory must have will also help constrain the search for new physical theories.
-\textcolor{blue}{[Can we make this paragraph more math-focused?]}\\
 
@@
 With formal training in computer science and engineering, he will additionally serve as the primary contact to interface with others in these disciplines who may be interested in the work.
-Aidala, Uribe, and Carcassi, and once identified, the Ph.D.~student, will generally meet weekly to discuss the project.
+Aidala, Uribe, and Carcassi, and once identified, the Ph.D.~student, will generally meet at least weekly to discuss the project.
 
@@
 Yet another possibility, since the JSD is the square of a distance function, is to identify sufficient conditions for a pseudo-distance to be the square of a distance, which may clarify what properties the entropic structure imposes.
-Developing skills in particular in information theory and information geometry will lead to a wide range of possible next steps in the student's career after their Ph.D., with applications for example in quantum computing and in machine learning across academia and the private sector.
+Developing proficiency in particular in information theory and information geometry will lead to a wide range of possible next steps in the student's career after their Ph.D., with applications for example in quantum computing and in machine learning across academia and the private sector.
 We additionally anticipate involving one or more undergraduates in the project, identifying self-contained pieces of the project to which they could contribute.
-\textcolor{blue}{For example, starting from a few specific results from information geometry or information theory, they could work on trying generalize these results to ensemble spaces.}
+For example, starting from a few specific results from information geometry or information theory, they could work on trying to generalize these results to ensemble spaces.
 
@@
 She served on the National Academy of Sciences Committee on a U.S.-Based Electron-Ion Collider Science Assessment 2016-18 and on the U.S.~Nuclear Science Advisory Committee 2023-25.
-She served as an elected member of the Executive Committee of the APS Division of Nuclear Physics (DNP) 2021-23, and since 2021 she has served as a DNP Ally to ensure that DNP meetings are welcoming and inclusive environments.
-Much of her work has been in quantum chromodynamics; however, in 2015 she initiated the Assumptions of Physics project with her collaborator and husband Gabriele Carcassi, and since 2018 they have so far published together 15 peer-reviewed articles [one more to add]~\cite{aop-phys-blueprint,aop-topExpDisting,aop-HamQuantInfo,Carcassi:2021,aop-spacetimeStruct,aop-HamConsInfoEnt,Carcassi2021four,Carcassi:2022bpm,aop-HamPriv,aop-commonLogical,aop-action,aop-nogo,aop-nonaddmeas,aop-unphysHilbert} as well as a technical book on the mathematical foundations of physics~\cite{aop-book}, with one additional article currently under review and publicly available as a preprint~\cite{aop-classicallimit}.
-She is the PI on a grant from the John Templeton Foundation that supports the development of novel measure theoretic structures for quantum mechanics.
-\\
+Much of her work has been in quantum chromodynamics; however, in 2015 she initiated the Assumptions of Physics program with her collaborator and husband Gabriele Carcassi, and since 2018 they have so far published together 15 peer-reviewed articles ~\cite{aop-phys-blueprint,aop-topExpDisting,aop-HamQuantInfo,Carcassi:2021,aop-spacetimeStruct,aop-HamConsInfoEnt,Carcassi2021four,Carcassi:2022bpm,aop-HamPriv,aop-commonLogical,aop-action,aop-nogo,aop-nonaddmeas,aop-nogoresponse,aop-unphysHilbert} as well as a technical book on the mathematical foundations of physics~\cite{aop-book}, with one additional article currently under review and publicly available as a preprint~\cite{aop-classicallimit}.
+She is the PI on a grant from the John Templeton Foundation that supports the development of novel measure-theoretic structures for quantum mechanics.
 
-\textcolor{red}{Alejandro: Add relevant qualifications.}
+Co-PI Uribe is a Fellow of the American Mathematical Society, with a long track record of publication in first-rate journals going back decades.
+He has held the positions of Professeur Invit\'{e} at the \'{E}cole Normale Sup\'{e}rieure (Paris) and at the Universit\'{e} de Paris-Saclay, and of Research Professor at MSRI (now the Simons Laufer Mathematical Sciences Institute).
+He was the PI on various NSF grants (DMS) from 1989 to 2014.
 
-Carcassi is the technical lead for the Assumptions of Physics project.
+Carcassi is the co-initiator with Aidala and technical lead for the Assumptions of Physics program.
 He is the Co-PI on the Templeton Foundation grant with Aidala.
-Related to the AoP project, he has given seminars to audiences of physicists, mathematicians, and philosophers at places including Harvard, Caltech, Oxford (UK) the University of Pavia (Italy) the Munich Center for Mathematical Philosophy (Germany) and the Wolfram Institute.
+Related to the AoP program, he has given seminars to audiences of physicists, mathematicians, and philosophers at places including Harvard, Caltech, Oxford (UK) the University of Pavia (Italy) the Munich Center for Mathematical Philosophy (Germany) and the Wolfram Institute.
 He and Aidala gave a joint public talk on misconceptions in the foundations of physics for the Michigan State University Advanced Studies Gateway, and he will give a U-M Saturday Morning Physics public talk on ``What is classical and what is quantum?'' in November 2025.
 He was invited and will deliver an overview talk on the AoP program at a special session of the Joint Mathematics Meeting in January 2026.
-His training in computer science and engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in physics, for example from information theory, theoretical computer science, systems theory, signal processing, and operations research.
-His position as U-M Physics Department staff without teaching responsibilities provides him with time and flexibility to dedicate to research related to the AoP project such as the work proposed here.
+His training in computer science and engineering allows him to bring ideas from fields not typically studied by individuals trained primarily in mathematics or physics, for example from information theory, theoretical computer science, systems theory, signal processing, and operations research.
+His position as U-M Physics Department staff without teaching responsibilities provides him with time and flexibility to dedicate to research related to the AoP program such as the work proposed here.
 See also his biographical sketch included as an Other Supplementary Document.\\
@@
 Together, Aidala, Uribe, and Carcassi are uniquely positioned to advance the mathematical foundations of physics, with Aidala's and Uribe's traditional training in physics and mathematics, respectively, complementing Carcassi's formal training in computer science and engineering.
-Carcassi's university staff position has allowed him to audit multiple graduate Mathematics and Physics courses at U-M, after having previously audited multiple graduate Physics courses at Stony Brook University.
-Uribe has served as a mathematical consultant to Aidala and Carcassi for the broader Assumptions of Physics project since 2018.
-For the AoP project, Aidala and Carcassi's published work so far has covered a wide range of topics.
+Carcassi's university staff position has allowed him to audit a number of graduate Mathematics and Physics courses at U-M, after having previously audited multiple graduate Physics courses at Stony Brook University.
+Uribe has served as a mathematical consultant to Aidala and Carcassi for the broader Assumptions of Physics program since 2018.
+For the AoP program, Aidala and Carcassi's published work so far has covered a wide range of topics.
 The ``physical mathematics'' approach within their work is an approach to the mathematical foundations of physics that seeks to construct mathematical structures strictly from axioms and definitions that can be rigorously justified from physical requirements~\cite{aop-book}.
@@
 
-Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award ``for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds,'' and in 2025 received the U-M Rackham Distinguished Graduate Mentor Award.
+Regarding success in mentoring and education, in 2019 Aidala received the U-M Imes and Moore Mentorship Award ``for exceptional contributions toward recruiting and mentoring graduate students in the sciences from disadvantaged and nontraditional backgrounds,'' and in 2025 the U-M Rackham Distinguished Graduate Mentor Award.
 She has served as the dissertation chair for 10 graduate students in Physics and Applied Physics who have already completed their doctorates, as co-chair for another in Applied Physics and as co-chair with Uribe for one in Applied and Indisciplinary Mathematics.
@@
 One of Carcassi's former undergraduate research mentees is now pursuing a Ph.D.
-in Mathematics at UT Austin.\\
+in Math at UT Austin.
 
-\textcolor{red}{Alejandro: Add info on mentoring.}
+Uribe has graduated six Ph.D.~students and has mentored 13 postdoctoral assistant professors.
+He has directed two Master's theses in the U-M Mathematics bridge program (MLB Scholars), as well as numerous REU projects.
+He has additionally served on the dissertation committee of four of Aidala's students thus far.
 
-Uribe has additionally served on the dissertation committee of four of Aidala's students thus far.
-
-Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with more than 17,800 subscribers currently.
+Regarding the qualifications of the research team to achieve the broader impacts beyond academia as described above, Carcassi has a YouTube channel (@gcarcassi) focused on physics and mathematics with 18k subscribers currently.
 He posts a new video monthly, generally targeted at individuals with an advanced undergraduate background in physics and/or mathematics.
-Collectively, his videos on this channel have received over 625k views.
+Collectively, his videos on this channel have received over 630k views.
 He additionally has a second, more technical YouTube channel (@AssumptionsofPhysicsResearch) aimed at professional researchers, with more than 1600 subscribers.
-He uses this channel as a venue to disseminate research results at a more technical level, including video versions of preprints or published papers as well as conference presentations related to the AoP project.
+He uses this channel as a venue to disseminate research results at a more technical level, including video versions of preprints or published papers as well as conference presentations related to the AoP program.
 He moreover uses the technical channel as an innovative approach to try to ``crowd-source'' technical expertise to advance the research more efficiently.
-Given the breadth of the broader AoP project, there is no single community with the appropriate technical background across all relevant and potentially relevant areas of mathematics and physics.
+Given the breadth of the broader AoP program, there is no single community with the appropriate technical background across all relevant and potentially relevant areas of mathematics and physics.
 So far this ``crowd-sourcing'' approach has led to multiple instances of viewers with whom we have had no previous contact pointing us to valuable reference materials, and in 2024 it led to our first coauthored paper with a researcher who reached out to us after seeing an open problem video, on ``Classical mechanics as the high-entropy limit of quantum mechanics''~\cite{aop-classicallimit}.
@@
 In other forms of reach beyond academia, Carcassi was interviewed in 2024 by Curt Jaimungal for the ``Theories of Everything'' podcast~\cite{Carcassi-ToEInterview}, which has more than 500k subscribers.
-Carcassi additionally writes monthly blog-style essays related to the AoP project and advertised through social media platforms such as Facebook and LinkedIn.
+Carcassi additionally writes monthly blog-style essays related to the AoP program and advertised through social media platforms such as Facebook and LinkedIn.
 In more creative outreach work, in April 2024 at U-M he performed an original musical comedy lecture, ``An Engineer in the Foundations of Physics,'' which drew approximately 100 attendees.
@@
 This proposal is the second time she is submitting to the Applied Mathematics program, based on the research direction in the mathematical foundations of physics that she initiated more recently, which has so far produced 15 peer-reviewed papers since 2018 as well as a technical book, as described above in the section on qualifications of the PI and research team.
-\\
-
-\textcolor{red}{Alejandro: Add information on intellectual merit of any prior NSF support.
-Technically only grants with an end date within the last five years are required to be listed, which I believe you don't have, but since I see you had several NSF grants between 1989 and ending in 2014, I think it would still be good to somehow discuss your prior record of NSF funding, perhaps in broader terms.}
+Uribe was the PI on various NSF grants from 1989--2014.
 
@@
 students and 7 undergraduates worked on the projects supported by Aidala's CAREER grant 2015-20.
-Among her former grad students, Bryan Ramson is now a permanent staff scientist at Fermi National Accelerator Lab, Joseph Osborn a permanent staff scientist at Brookhaven National Lab, Catherine Ayuso Head of Special Projects and Research Manager at Entanglement, Inc., Dylan Manna the owner of Manna Management Consulting, and Ibrahim Chahrour a lecturer at U-M.
-Others are in postdoctoral positions or still students in the group.
-\\
-
-\textcolor{red}{Alejandro: Add broader impacts of any prior NSF support.}
+Her former grad students have gone on to permanent staff scientist positions at Fermi National Accelerator Lab and Brookhaven National Lab, a faculty position at Nantes Universit\'{e} (France), a Lecturer position at U-M, the Head of Special Projects and Research Manager at Entanglement, Inc., and to found their own consulting firm, with others in postdoctoral positions.
 

### 2025-11-08 Proposals/2025-NSF-EntropicGeometry/ProjectSummary.tex
Further edits after discussion with G.
Various further updates

--- 2025-09-22 Proposals/2025-NSF-EntropicGeometry/ProjectSummary.tex
@@
 \thispagestyle{empty}
+
 \noindent
 \textbf{Overview:}\\
+\textcolor{blue}{The goal of the proposed work is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum.
+The immediate aim is to generalize results of both classical and quantum information theory to a common, more abstract theory that can also handle the infinite-dimensional case.
+It takes the notion of statistical ensemble as a fundamental concept, as any physical theory must \emph{at least} provide us with a space of statistical ensembles.
+The relationships among the four mathematical structures required by a space of ensembles---the topology, the convex structure, the proposed entropic structure and the Lie algebra---will be investigated.} \\
 
@@
 \textbf{Intellectual Merit:}  \\
+\textcolor{blue}{The primary advancement will be to identify and characterize a set of mathematical structures common to \emph{any} physical theory, explicitly showing how these general structures reduce to the standard structures in the cases of discrete classical, continuous classical, and quantum systems.
+Such a generalization will provide further insight into the relationships among established physical theories and place constraints on the allowed mathematical structures in potential future physical theories.
+All axioms will be physically motivated and stated as properties of ensemble spaces.}
+
+\textcolor{blue}{The proposed work starts from the observation that the spaces of statistical ensembles of classical and quantum mechanics have multiple features in common.
+They are both convex subsets of a locally convex topological vector space whose metric is induced by the information entropy, or more specifically, by the Jensen-Shannon divergence.
+This is in contrast to the different structures of the spaces of classical and quantum states, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces, respectively.
+Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can indeed be recovered by the entropy.
+In general, the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
+While classical mechanics and quantum mechanics have a different structure, classical and quantum statistical mechanics have common features.
+Moreover, classical and quantum information theory have numerous corresponding features.
+Our investigation will take those common features and generalize them, providing a common foundational structure to both classical and quantum theories.
+}
 
 \noindent
-\textbf{Broader impacts:}  \\
-
-The proposed activities additionally include significant outreach through social media, primarily through YouTube.
-Videos will be produced to disseminate the results of the research, with some videos aimed at the level of an advanced undergraduate in mathematics and/or physics, furthering education.
+\textbf{Broader Impacts:} \\
+The proposed activities include an annual online summer school and significant outreach through social media, primarily through YouTube.
+The summer school will take place over three days, with six hours of lectures and three hours of discussion total, presenting ideas and results from the project at a level appropriate to advanced undergraduates and above.
+For the outreach through social media, videos will be produced to disseminate the results of the research, with some videos aimed at the level of an advanced undergraduate in mathematics and/or physics.
 Other videos will be aimed at active researchers in various fields of mathematics and physics, providing a new mathematical framework as well as a new perspective linking traditionally disparate realms of physics that may impact tools or thinking in their own research.

### 2025-11-08 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Final version of project summary, significance, and bibliography formatting before talking to Alejandro

@@
 This gives the ensemble space a convex structure as $p \ens[a] + \bar{p} \ens[b] \in \Ens$ for all $\ens[a], \ens[b] \in \Ens$ and $p \in [0,1]$.
-This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs); see e.g.~Refs.~\cite{plavala2023,wilce2025generalizedprobabilitytheorynotes,DArianoBook} \textcolor{blue}{Anything else to cite?}.
+This insight is already used in the literature, particularly in Generalized Probabilistic Theories (GPTs); see e.g.~Refs.~\cite{plavala2023,wilce2025generalizedprobabilitytheorynotes,DArianoBook}.
 The main innovation in this respect is that we do not limit ourselves to finite combinations and we do not simply allow all infinite combinations (which can be shown to include physically untenable cases~\cite{baccetti2013infinite,aop-unphysHilbert}).
@@
 \textbf{Expected outcomes and significance:}
-During the three-year period of the proposed research, we expect to achieve \textcolor{blue}{and submit manuscripts on} at least the following:
+During the three-year period of the proposed research, we expect to achieve and submit manuscripts on at least the following:
 \begin{itemize}
@@
 
-\textcolor{blue}{The immediate significance of the proposed research stems from the realization that, though classical information theory and quantum information theory are presented as two separate subjects in textbooks, there are a lot of commonalities that make them an interesting target for generalization.
-The broader significance lies in its approach as well as its specific potential to provide a common axiomatic foundation for \emph{all} physical theories.
+\textcolor{blue}{The immediate significance of the proposed research stems from the realization that, though classical information theory and quantum information theory are presented as two separate subjects in textbooks, their numerous corresponding features make them an interesting target for generalization.
+The broader significance lies in the overall approach of the research as well as its specific potential to provide a common axiomatic foundation for \emph{all} physical theories.
 Currently within physics, we typically do not know from where the various mathematical structures emerge, making it unclear in which directions they can be generalized in order to develop new theories that are physically meaningful.
 Our approach instead seeks to \emph{find axioms from physical requirements}, i.e.~the axiomatics are justified from the physics.
-Such an axiomatic foundation common to all physical theories would serve not only to better ground current theories and elucidate their relationships, but also to constrain potential future physical theories.}
+Such an axiomatic foundation common to all physical theories would serve not only to better ground current theories and elucidate their relationships, but also to constrain potential future physical theories.} While there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature attempting to relate the geometric and entropic structures of either classical or quantum mechanics, much less those of a general physical theory.
+Works trying to reconstruct quantum mechancis from physical premises are typically in finite dimensions, do not generalize to classical mechanics, and do not provide a mathematically rigorous path to generalize to classical or quantum field theories.
 
 %We're checking that the axioms are independent of each other.
-
-While there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature attempting to relate the geometric and entropic structures of either classical or quantum mechanics, much less those of a general physical theory.
-Works trying to reconstruct quantum mechancis from physical premises are typically in finite dimensions, do not generalize to classical mechanics, and do not provide a mathematically rigorous path to generalize to classical or quantum field theories.
 

### 2025-11-08 Proposals/2025-NSF-EntropicGeometry/ProjectSummary.tex
Final version of project summary, significance, and bibliography formatting before talking to Alejandro

@@
 \textbf{Intellectual Merit:}  \\
-\textcolor{blue}{The primary advancement will be to identify and characterize a set of mathematical structures common to \emph{any} physical theory, explicitly showing how these general structures reduce to the standard structures in the cases of discrete classical, continuous classical, and quantum systems.
+\textcolor{blue}{While the state spaces of classical mechanics and quantum mechanics have different structures, symplectic manifolds and projective spaces of complex Hilbert spaces, respectively, the spaces of classical and quantum \emph{statistical ensembles} have common features.
+They are both convex subsets of a locally convex topological vector space whose metric is induced by the information entropy, or more specifically, by the Jensen-Shannon divergence.
+Moreover, classical and quantum information theory have numerous corresponding features.
+The primary advancement of the research will be to take those common and corresponding features and generalize them, providing a common, mathematically rigorous, foundational structure to both classical and quantum theories.
 Such a generalization will provide further insight into the relationships among established physical theories and place constraints on the allowed mathematical structures in potential future physical theories.
-All axioms will be physically motivated and stated as properties of ensemble spaces.}
+All axioms will be physically motivated and stated as properties of ensemble spaces.
+Preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can be recovered by the entropy, tightly linking the geometric and entropic structures, and that the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
+}
 
-\textcolor{blue}{The proposed work starts from the observation that the spaces of statistical ensembles of classical and quantum mechanics have multiple features in common.
-They are both convex subsets of a locally convex topological vector space whose metric is induced by the information entropy, or more specifically, by the Jensen-Shannon divergence.
-This is in contrast to the different structures of the spaces of classical and quantum states, which rely on symplectic manifolds and projective spaces of complex Hilbert spaces, respectively.
-Even more notably, our preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can indeed be recovered by the entropy.
-In general, the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
-While classical mechanics and quantum mechanics have a different structure, classical and quantum statistical mechanics have common features.
-Moreover, classical and quantum information theory have numerous corresponding features.
-Our investigation will take those common features and generalize them, providing a common foundational structure to both classical and quantum theories.
+\textcolor{blue}{
 }

### 2025-11-08 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
2025NSF: integrated most comments from Alejandro

@@
 The JSD is, in both cases, the square of a distance function and can be interpreted as how much the average entropy increases under an equal mixture of two ensembles.
-There are, therefore, three common structures at play: a topology (which physically represents experimental verifiability), a convex structure (which physically represents the ability to make statistical mixtures) and an entropic structure (which physically represents the variability of the elements within each ensemble).
-While the axioms for the topology and the convex structure are readily available, to our knowledge, the correct axioms for the entropic structure still need to be identified.
-Moreover, the interplay between the topology, the convex structure and the entropic structure opens additional mathematical questions which have not been investigated.
+There are, therefore, three common structures at play: a topology (which physically represents experimental verifiability), a convex structure (which physically represents the ability to make statistical mixtures) and an entropy function (which physically represents the variability of the elements within each ensemble).
+While the axioms for the topology and for convex spaces(CITE) are readily available, to our knowledge, the correct axioms for an entropy function and its required compatibility condition with the other two structure still need to be identified.
 
-The goal of this proposed work is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum.
+The goal of this proposed work is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum, which we will call \textbf{entropic structure}.
 The immediate aim is to generalize results of both classical and quantum information theory to a common, more abstract theory that can also handle the infinite-dimensional case.
@@
 
+Before looking at our preliminary proposal for the axiomatic, let us develop some intuition about the structure we need to generalize.
 Note that in all above three cases the definition of the entropy is tightly bound to the definition of the space.
 Therefore, the biggest problem is understanding how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each case.
-Fortunately, in our preliminary work we already have found the solution.
+Fortunately, in our preliminary work we have found the right approach.
 In all cases, the entropy possesses an upper bound to its increase under convex combination: $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$ where $\bar{p} = 1-p$ and $I(p) = -p \log p - \bar{p}\log \bar{p}$.
-This maximum increase is given precisely when the two ensembles are orthogonal with respect to the inner product of the respective space.
-In the classical case, this means that the measures have disjoint support; in the quantum case, it means that the eigenspaces of non-zero eigenvalues are orthogonal.
-In other words, characterizing the interaction of the entropy function with the topology and the convex structure is enough to characterize the entropic structure.
-That is, the form of $I(p)$---the Shannon entropy formula~\cite{Shannon}---can be characterized, and since in both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case.
+This maximum increase is given, in the classical case, when the two measures have disjoint support; in the quantum case, it is given when the eigenspaces of non-zero eigenvalues are orthogonal.
+Our approach, then, is to proceed in the opposite way: impose a generic entropy function to have an upper bound and enough properties to show that said upper bound must correspond to the Shannon entropy formula~\cite{Shannon}.
+Since both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case as a recursive expansion of the upper bound.
 
 Another important insight from our preliminary work is that the entropy is not only responsible for the geometry of the ensemble space, but it is also intimately related to the geometry of the state space of both classical and quantum mechanics.
-That is, the entropic structure of the ensemble space is equivalent to the geometric structure of the state space.
+The idea, then, is to recover to the geometric structure of the state space from it entropic structure.
 
-As we saw in classical mechanics, the symplectic form gives us the Liouville measure which in turn is used to define the entropy.
+For example, as we saw in classical mechanics, the symplectic form gives us the Liouville measure which in turn is used to define the entropy.
 Therefore, the entropy of an ensemble is defined from the geometry of the state space.
@@
 \end{equation}
+as this is the cosine square of the distance function of the Fubini-Study metric.
 Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has von Neumann entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi|\psi)}}{2} \right)$ where $I$, as before, is the Shannon entropy for two elements.
@@
 However, the relationship between $S(\rho_{\psi\phi})$ and $p(\phi|\psi)$ is invertible.
-This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the geometry of the state space.
+This means that, given the entropy of the equal mixtures of all pairs of states we can reconstruct the Born rule and, therefore, the geometry.
 
@@
 This provides us with a strong conceptual starting point, and properties of ensembles are what need to be specified through axioms.
+We will therefore call \textbf{ensemble space} a set endowed with the suitable topology, convex structure and entropic structure that generalizes the spaces of classical and quantum ensembles.
 
@@
 That is, infinite convex combinations exist only if they are the topological limit of finite convex combinations.
-On physical grounds, the mixing operation must be continuous in the natural topology defined by experimental verifiability, much like addition and scalar multiplication are continuous in a topological vector space, leading to the idea of a topological convex space.
+On physical grounds, the mixing operation must be continuous in the natural topology defined by experimental verifiability, much like addition and scalar multiplication are continuous in a topological vector space.
+Given the analogy, we call \textbf{topological convex space} a convex space that is also a topological space for which the mixing operation is continuous.
 
-Surprisingly, topological convex spaces do not seem to have been studied, and this structure alone presents interesting sets of questions.
+Topological convex spaces have not been developed in the literature, and this structure alone presents interesting sets of questions.
 For example, it is established that a convex space embeds into a vector space if and only if it is cancellative, that is, if $p \ens[a] + \bar{p} \ens = p \ens[b] + \bar{p} \ens$ for some $p \in (0,1)$ implies $\ens[a] = \ens[b]$.
@@
 
-In the finite-dimensional case, and if the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
-Let $V_{\ens[a]}$ be the embedding vector space with $\ens[a] \in \Ens$ as origin.
-This will be a differentiable manifold.
+Given an internal point $\ens[a] \in \Ens$ (i.e.
+for every $\ens[b] \in \Ens$ we can find $\ens[c] \in \Ens$ such that $\ens[a] = p \ens[b] + \bar{p} \ens[c]$ from some $p \in (0,1)$) we can construct an embedding vector space $V_{\ens[a]}$.
+If $V_{\ens[a]}$ is finite dimensional, we can show that $V_{\ens[a]}$ is a neighborhood of $\ens[a]$, which means that the interior of $\Ens$ is a manifold.
+If the entropy is twice differentiable with respect to the convex coefficients, the strict concavity of the entropy means that the Hessian is negative definite, which allows us to define a metric tensor.
 On a linear coordinate chart we can define $g_{\ens}( v, w ) = - \frac{\partial^2 S}{\partial \ens ^2}( v, w )$ where $v$ and $w$ are elements of the tangent space at $\ens \in \Ens \subset V_{\ens[a]}$.
@@
 
-Another goal is to recover the geometric structure of the state spaces of both classical and quantum mechanics from the entropic structure, to formally show the equivalence.
-For quantum mechanics, one can simply define the Born rule from the entropy as we have seen before.
-For classical mechanics, the situation is a bit more complex as we need to recover first the Liouville measure.
+Another construction allows us to generalize the notion of count of states given, in classical mechanics, by the Liouville measure.
 This can be done by noting two things: for a uniform distribution with support $U \subseteq X$, $\mu(U) = 2^{S(\rho_U)}$, and the uniform distribution over $U$ has the highest entropy of all distributions with support $U$.
-Given a set $A \subseteq \Ens$ of ensembles, we can then define the state capacity as the supremum of the exponential of the hull of $A$, which physically means the highest entropy achievable with a statistical mixture of elements of $A$.
+Given a set $A \subseteq \Ens$ of ensembles, we can then define the state capacity as the supremum of the exponential of the entropy within the hull of $A$, which physically means the highest entropy achievable with a statistical mixture of elements of $A$.
 This set function has nice properties: it is non-negative, monotonic, subadditive in general and additive on orthogonal sets of ensembles.

### 2025-11-08 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
2025NSF: better flow and removed repetitions

@@
 
-Before looking at our preliminary proposal for the axiomatic, let us develop some intuition about the structure we need to generalize.
-Note that in all above three cases the definition of the entropy is tightly bound to the definition of the space.
-Therefore, the biggest problem is understanding how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each case.
-Fortunately, in our preliminary work we have found the right approach.
-In all cases, the entropy possesses an upper bound to its increase under convex combination: $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$ where $\bar{p} = 1-p$ and $I(p) = -p \log p - \bar{p}\log \bar{p}$.
-This maximum increase is given, in the classical case, when the two measures have disjoint support; in the quantum case, it is given when the eigenspaces of non-zero eigenvalues are orthogonal.
-Our approach, then, is to proceed in the opposite way: impose a generic entropy function to have an upper bound and enough properties to show that said upper bound must correspond to the Shannon entropy formula~\cite{Shannon}.
-Since both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case as a recursive expansion of the upper bound.
+The central focus is the intuitive notion of statistical ensemble (i.e.~the collection of all outputs of a preparation procedure), as this seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
+In terms of the mathematics, as we have seen, the spaces of statistical ensembles of classical and quantum mechanics can be made to look very similar.
+From the experimental side, statistical ensembles are what we produce and measure experimentally.
+On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form ``whenever I prepare a system according to this procedure, I will find this range of outcomes.'' That is, physical laws are relationships between ensembles.
+Therefore, it makes sense to require that any physical theory must \emph{at least} posit a space of ensembles.
+This provides us with a strong conceptual starting point, and properties of ensembles are what need to be specified through axioms.
+We will therefore call \textbf{ensemble space} a set endowed with the suitable topology, convex structure and entropic structure that generalizes the spaces of classical and quantum ensembles.
 
-Another important insight from our preliminary work is that the entropy is not only responsible for the geometry of the ensemble space, but it is also intimately related to the geometry of the state space of both classical and quantum mechanics.
-The idea, then, is to recover to the geometric structure of the state space from it entropic structure.
+Before proceeding, we should not that the entropic structure is not crucial for the geometry of the ensemble space, but it is also intimately related to the geometry of the state space of both classical and quantum mechanics.
+One direction we want to pursue, in fact, is to recover to the geometric structure of the state space from it entropic structure.
 
@@
 
-Given that \emph{all} physical theories must admit a notion of ensembles that can be statistically mixed via convex combination, these insights from our preliminary work show that the entropic structure plays a crucial role in all physical theories and merits further study.
-Moreover, while typically information geometry is cast in finite-dimensional spaces, there is nothing preventing its generalization to infinite dimensions.
-If the entropic structure, in fact, imposes a distance function whose open balls are convex, then this provides us with a locally convex topology which is exactly what is needed to define calculus in infinite dimensions.
-Moreover, the Lie algebras of both classical and quantum mechanics, Poisson brackets and commutators, respectively, have a special relationship with the geometric and entropic structures, which may provide us with a template to generalize those structures as well.
+Lastly, as port of our larger program, we have shown that classical mechanics emerges as the high entropy limit of quantum mechanics.
+These insights show that the entropic structure is more fundamental that one may expect, and that finding its axiomatic characterization should allow for a more precise generalization of a series of results.
+We stress that the axioms that characterize the entropic structure of physical theories should be minimal, in the sense that, in the spirit of analogous work on the foundations of mathematics, like set theory or reverse mathematics~\cite{RevMathBook}, the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory.
+Part of the work, then, is to understand the implication relationships between different requirements.
+Lastly, note that the Lie algebras of both classical and quantum mechanics, Poisson brackets and commutators, respectively, have a special relationship with the geometric and entropic structures, which may provide us with a template to generalize those structures as well.
 
-Given these very promising preliminary findings, the goal of the proposed research is to find a minimal set of axioms to characterize the entropic structure of physical theories.
-Minimal in the sense that, in the spirit of analogous work on the foundations of mathematics, like set theory or reverse mathematics~\cite{RevMathBook}, the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory.
-Let us, then, go through the current state of our preliminary work, and show the type of questions that the overall approach touches.
+Before delving into the axiomatic, note that in all above three cases the definition of the entropy is tightly bound to the definition of the space.
+Therefore, we need to understand how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each case.
+Fortunately, in our preliminary work we have found the right approach.
+In all cases, the entropy possesses an upper bound to its increase under convex combination: $S(p\ens[a] + \bar{p} \ens[b]) \leq I(p) + p S(\ens[a]) + \bar{p} S(\ens[b])$ where $\bar{p} = 1-p$ and $I(p) = -p \log p - \bar{p}\log \bar{p}$.
+This maximum increase is given, in the classical case, when the two measures have disjoint support; in the quantum case, it is given when the eigenspaces of non-zero eigenvalues are orthogonal.
+Our approach, then, is to proceed in the opposite way: impose a generic entropy function to have an upper bound and enough properties to show that said upper bound must correspond to the Shannon entropy formula~\cite{Shannon}.
+Since both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case as a recursive expansion of the upper bound.
+
+Having justified why studying the entropic structure of ensemble spaces is of interest and having found a fruitful approach, let us go through the axioms of each structure from our preliminary work.
 
@@
 Ultimately, answering these questions means understanding both the physical and mathematical consequences of those decisions.
-
-The central focus is the intuitive notion of statistical ensemble (i.e.~the collection of all outputs of a preparation procedure), as this seems to be particularly suitable for generalization for a number of both practical and conceptual considerations.
-In terms of the mathematics, as we have seen, the spaces of statistical ensembles of classical and quantum mechanics can be made to look very similar.
-From the experimental side, statistical ensembles are what we produce and measure experimentally.
-On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form ``whenever I prepare a system according to this procedure, I will find this range of outcomes.'' That is, physical laws are relationships between ensembles.
-Therefore, it makes sense to require that any physical theory must \emph{at least} posit a space of ensembles.
-This provides us with a strong conceptual starting point, and properties of ensembles are what need to be specified through axioms.
-We will therefore call \textbf{ensemble space} a set endowed with the suitable topology, convex structure and entropic structure that generalizes the spaces of classical and quantum ensembles.
 

### 2025-11-09 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Fix typos, add refs, remove blue text

@@
 There are, therefore, three common structures at play: a topology (which physically represents experimental verifiability), a convex structure (which physically represents the ability to make statistical mixtures) and an entropy function (which physically represents the variability of the elements within each ensemble).
-While the axioms for the topology and for convex spaces(CITE) are readily available, to our knowledge, the correct axioms for an entropy function and its required compatibility condition with the other two structure still need to be identified.
+While the axioms for the topology and for convex spaces~\cite{Gudder1973,fritz2015convexspaces} are readily available, to our knowledge, the correct axioms for an entropy function and its required compatibility condition with the other two structures still need to be identified.
 
@@
 From the experimental side, statistical ensembles are what we produce and measure experimentally.
-On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form ``whenever I prepare a system according to this procedure, I will find this range of outcomes.'' That is, physical laws are relationships between ensembles.
+On the conceptual side, the reproducibility requirement of scientific investigation means that physical laws are of the form, ``Whenever I prepare a system according to this procedure, I will find this range of outcomes.'' That is, physical laws are relationships between ensembles.
 Therefore, it makes sense to require that any physical theory must \emph{at least} posit a space of ensembles.
@@
 
-Before proceeding, we should not that the entropic structure is not crucial for the geometry of the ensemble space, but it is also intimately related to the geometry of the state space of both classical and quantum mechanics.
-One direction we want to pursue, in fact, is to recover to the geometric structure of the state space from it entropic structure.
+Before proceeding, we should note that the entropic structure is not only crucial for the geometry of the ensemble space, but it is also intimately related to the geometry of the state space of both classical and quantum mechanics.
+One direction we want to pursue, in fact, is to recover the geometric structure of the state space from its entropic structure.
 
-For example, as we saw in classical mechanics, the symplectic form gives us the Liouville measure which in turn is used to define the entropy.
+For example, as we saw in classical mechanics, the symplectic form gives us the Liouville measure, which in turn is used to define the entropy.
 Therefore, the entropy of an ensemble is defined from the geometry of the state space.
@@
 \begin{equation}
-	p(\phi|\psi) = \frac{\< \phi | \psi \>\< \psi | \phi \>}{\< \psi | \psi \>\< \phi | \phi \>}.
+	p(\phi|\psi) = \frac{\< \phi | \psi \>\< \psi | \phi \>}{\< \psi | \psi \>\< \phi | \phi \>},
 \end{equation}
-as this is the cosine square of the distance function of the Fubini-Study metric.
+as this is the cosine squared of the distance function of the Fubini-Study metric.
 Given two states $\psi$ and $\phi$, the ensemble given by their equal mixture, $\rho_{\psi\phi} = \frac{1}{2}|\psi\>\<\psi|+\frac{1}{2}|\phi\>\<\phi|$,  has von Neumann entropy $S(\rho_{\psi\phi}) = I\left(\frac{1 + \sqrt{p(\phi|\psi)}}{2} \right)$ where $I$, as before, is the Shannon entropy for two elements.
@@
 
-Lastly, as port of our larger program, we have shown that classical mechanics emerges as the high entropy limit of quantum mechanics.
-These insights show that the entropic structure is more fundamental that one may expect, and that finding its axiomatic characterization should allow for a more precise generalization of a series of results.
+Lastly, as part of our larger program, we have shown that classical mechanics emerges as the high-entropy limit of quantum mechanics~\cite{aop-classicallimit}.
+These insights show that the entropic structure is more fundamental than one may expect, and that finding its axiomatic characterization should allow for a more precise generalization of a series of results.
 We stress that the axioms that characterize the entropic structure of physical theories should be minimal, in the sense that, in the spirit of analogous work on the foundations of mathematics, like set theory or reverse mathematics~\cite{RevMathBook}, the axioms can be shown to be satisfied by the different theories (which now become models of the larger theory) but also allow one to rederive the entropy formula for each specific theory.
@@
 
-Before delving into the axiomatic, note that in all above three cases the definition of the entropy is tightly bound to the definition of the space.
+Before delving into the axiomatics, note that in all three cases above the definition of the entropy is tightly bound to the definition of the space.
 Therefore, we need to understand how one can give a single characterization that is general enough to encompass the three cases, while constrained enough to recover the correct expression in each case.
@@
 Our approach, then, is to proceed in the opposite way: impose a generic entropy function to have an upper bound and enough properties to show that said upper bound must correspond to the Shannon entropy formula~\cite{Shannon}.
-Since both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case as a recursive expansion of the upper bound.
+Since in both classical and quantum mechanics ensembles can always be decomposed into orthogonal components, the different entropy formulas can be recovered in each case as a recursive expansion of the upper bound.
 
@@
 
-The second axiom imposes a convex structure, which represents the ability to perform statistical mixtures.
+The second axiom imposes a convex structure~\cite{Gudder1973,fritz2015convexspaces}, which represents the ability to perform statistical mixtures.
 Given two ensembles $\ens[a], \ens[b] \in \Ens$, we can always choose a probability $p \in [0,1]$ and select $\ens[a]$ $p$ percent of the times and $\ens[b]$ $\bar{p} = 1-p$ percent of the times.
@@
 Topological convex spaces have not been developed in the literature, and this structure alone presents interesting sets of questions.
-For example, it is established that a convex space embeds into a vector space if and only if it is cancellative, that is, if $p \ens[a] + \bar{p} \ens = p \ens[b] + \bar{p} \ens$ for some $p \in (0,1)$ implies $\ens[a] = \ens[b]$.
+For example, it is established that a convex space embeds into a vector space if and only if it is cancellative~\cite{CapraroFritz2013}, that is, if $p \ens[a] + \bar{p} \ens = p \ens[b] + \bar{p} \ens$ for some $p \in (0,1)$ implies $\ens[a] = \ens[b]$.
 The question now is what are necessary and sufficient conditions for a topological convex space to embed continuously into a topological vector space?
@@
 Moreover, in the case of both classical and quantum mechanics, states can be decomposed into orthogonal components, which means the upper variability is an equality and can be used to calculate the entropy from that decomposition.
-This will allows us to recover the expressions for the entropy in the three different cases, classical discete, classical continuous, and quantum.
-Similarly, we will be able to show that, from this structure alone, the geometric structure of both classical and quantum mechanics can be recovered.
+This will allow us to recover the expressions for the entropy in the three different cases, classical discrete, classical continuous, and quantum.
+Similarly, we will be able to show that, from this structure alone, the geometric structures of both classical and quantum mechanics can be recovered.
 
-Our preliminary work also shows that an entropy satisfying those properties forces the convex space to be cancellative, meaning that it embeds into a vector space.
-That is, a convex space that does not embed into a vector space cannot posses an entropic structure.
+Our preliminary work also shows that an entropy satisfying those properties forces the convex space to be cancellative, meaning that it embeds into a vector space~\cite{CapraroFritz2013}.
+That is, a convex space that does not embed into a vector space cannot possess an entropic structure.
 We can, in fact, produce a model of that case and see exactly why the entropic structure fails.
@@
 
-Given an internal point $\ens[a] \in \Ens$ (i.e.
-for every $\ens[b] \in \Ens$ we can find $\ens[c] \in \Ens$ such that $\ens[a] = p \ens[b] + \bar{p} \ens[c]$ from some $p \in (0,1)$) we can construct an embedding vector space $V_{\ens[a]}$.
+Given an internal point $\ens[a] \in \Ens$ (i.e.~for every $\ens[b] \in \Ens$ we can find $\ens[c] \in \Ens$ such that $\ens[a] = p \ens[b] + \bar{p} \ens[c]$ from some $p \in (0,1)$), we can construct an embedding vector space $V_{\ens[a]}$.
 If $V_{\ens[a]}$ is finite dimensional, we can show that $V_{\ens[a]}$ is a neighborhood of $\ens[a]$, which means that the interior of $\Ens$ is a manifold.
@@
 
-Another construction allows us to generalize the notion of count of states given, in classical mechanics, by the Liouville measure.
-This can be done by noting two things: for a uniform distribution with support $U \subseteq X$, $\mu(U) = 2^{S(\rho_U)}$, and the uniform distribution over $U$ has the highest entropy of all distributions with support $U$.
-Given a set $A \subseteq \Ens$ of ensembles, we can then define the state capacity as the supremum of the exponential of the entropy within the hull of $A$, which physically means the highest entropy achievable with a statistical mixture of elements of $A$.
+Another construction allows us to generalize the notion of count of states, given in classical mechanics by the Liouville measure.
+This can be done by noting two things: for a uniform distribution over the state space $X$ with support $U \subseteq X$, $\mu(U) = 2^{S(\rho_U)}$, and the uniform distribution over $U$ has the highest entropy of all distributions with support $U$.
+Given a set $A \subseteq \Ens$ of ensembles, we can then define the \textbf{state capacity} as the supremum of the exponential of the entropy within the hull of $A$, which physically means the highest entropy achievable with a statistical mixture of elements of $A$.
 This set function has nice properties: it is non-negative, monotonic, subadditive in general and additive on orthogonal sets of ensembles.
@@
 
-\textcolor{blue}{The immediate significance of the proposed research stems from the realization that, though classical information theory and quantum information theory are presented as two separate subjects in textbooks, their numerous corresponding features make them an interesting target for generalization.
+The immediate significance of the proposed research stems from the realization that, though classical information theory and quantum information theory are presented as two separate subjects in textbooks, their numerous corresponding features make them an interesting target for generalization.
 The broader significance lies in the overall approach of the research as well as its specific potential to provide a common axiomatic foundation for \emph{all} physical theories.
@@
 Our approach instead seeks to \emph{find axioms from physical requirements}, i.e.~the axiomatics are justified from the physics.
-Such an axiomatic foundation common to all physical theories would serve not only to better ground current theories and elucidate their relationships, but also to constrain potential future physical theories.} While there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature attempting to relate the geometric and entropic structures of either classical or quantum mechanics, much less those of a general physical theory.
-Works trying to reconstruct quantum mechancis from physical premises are typically in finite dimensions, do not generalize to classical mechanics, and do not provide a mathematically rigorous path to generalize to classical or quantum field theories.
+Such an axiomatic foundation common to all physical theories would serve not only to better ground current theories and elucidate their relationships, but also to constrain potential future physical theories.
+While there exists other work attempting to reconstruct quantum mechanics from physical principles, for example in Refs.~\cite{hardy2001quantumaxioms,caticha2008,dakic2009,Chiribella2010,Masanes_2011}, we are not aware of any literature attempting to relate the geometric and entropic structures of either classical or quantum mechanics, much less those of a general physical theory.
+Works trying to reconstruct quantum mechanics from physical premises are typically in finite dimensions, do not generalize to classical mechanics, and do not provide a mathematically rigorous path to generalize to classical or quantum field theories.
 
@@
 \noindent
-\textbf{Relationship to other current funding:} PI Aidala's current funding from the NSF Division of Physics is for work on LHCb experiment at CERN, and from the NSF AccelNet program is to increase collaboration on quantum chromodynamics (QCD) research between the U.S.~and Latin America.
+\textbf{Relationship to other current funding:} PI Aidala's current funding from the NSF Division of Physics is for work on the LHCb experiment at CERN, and from the NSF AccelNet program is to increase collaboration on quantum chromodynamics (QCD) research between the U.S.~and Latin America.
 The PI's funding from the Department of Energy is for work at the Relativistic Heavy Ion Collider (RHIC) on the PHENIX and sPHENIX experiments.

### 2025-11-09 Proposals/2025-NSF-EntropicGeometry/ProjectSummary.tex
Fix typos, add refs, remove blue text

@@
 \textbf{Overview:}\\
-\textcolor{blue}{The goal of the proposed work is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum.
+The goal of the proposed work is to develop a common axiomatic basis for the information-theoretic structure underlying all physical theories, classical and quantum.
 The immediate aim is to generalize results of both classical and quantum information theory to a common, more abstract theory that can also handle the infinite-dimensional case.
 It takes the notion of statistical ensemble as a fundamental concept, as any physical theory must \emph{at least} provide us with a space of statistical ensembles.
-The relationships among the four mathematical structures required by a space of ensembles---the topology, the convex structure, the proposed entropic structure and the Lie algebra---will be investigated.} \\
+The relationships among the four mathematical structures required by a space of ensembles---the topology, the convex structure, the proposed entropic structure and the Lie algebra---will be investigated.
+\\
 
@@
 \textbf{Intellectual Merit:}  \\
-\textcolor{blue}{While the state spaces of classical mechanics and quantum mechanics have different structures, symplectic manifolds and projective spaces of complex Hilbert spaces, respectively, the spaces of classical and quantum \emph{statistical ensembles} have common features.
+While the state spaces of classical mechanics and quantum mechanics have different structures, symplectic manifolds and projective spaces of complex Hilbert spaces, respectively, the spaces of classical and quantum \emph{statistical ensembles} have common features.
 They are both convex subsets of a locally convex topological vector space whose metric is induced by the information entropy, or more specifically, by the Jensen-Shannon divergence.
@@
 Preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can be recovered by the entropy, tightly linking the geometric and entropic structures, and that the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
-}
 

### 2025-11-09 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Final read of broader impacts

@@
 For example, starting from a few specific results from information geometry or information theory, they could work on trying to generalize these results to ensemble spaces.
+No funding is requested for undergraduates; there are various mechanisms for undergraduates to engage in research such as research for credit, summer funding to which U-M students can apply, and competitive programs bringing external students to U-M over the summer for paid research.
 
@@
 Given the breadth of the broader AoP program, there is no single community with the appropriate technical background across all relevant and potentially relevant areas of mathematics and physics.
-So far this ``crowd-sourcing'' approach has led to multiple instances of viewers with whom we have had no previous contact pointing us to valuable reference materials, and in 2024 it led to our first coauthored paper with a researcher who reached out to us after seeing an open problem video, on ``Classical mechanics as the high-entropy limit of quantum mechanics''~\cite{aop-classicallimit}.
-We were previously completely unaware of the researcher and his work in Innsbruck, Austria.
-In other forms of reach beyond academia, Carcassi was interviewed in 2024 by Curt Jaimungal for the ``Theories of Everything'' podcast~\cite{Carcassi-ToEInterview}, which has more than 500k subscribers.
+So far this ``crowd-sourcing'' approach has led to multiple instances of viewers with whom we have had no previous contact pointing us to valuable reference materials, and in 2024 it led to our first coauthored paper with a researcher who reached out to us from Innsbruck, Austria after seeing an open problem video, on ``Classical mechanics as the high-entropy limit of quantum mechanics''~\cite{aop-classicallimit}.
+In other forms of reach beyond academia, Carcassi was interviewed in 2024 by Curt Jaimungal for the well-known ``Theories of Everything'' podcast~\cite{Carcassi-ToEInterview}, which has more than 500k subscribers; interviewees in the prior and subsequent weeks included Roger Penrose and Leonard Susskind.
 Carcassi additionally writes monthly blog-style essays related to the AoP program and advertised through social media platforms such as Facebook and LinkedIn.

### 2025-11-10 Proposals/2025-NSF-EntropicGeometry/ProjectSummary.tex
Change Project Summary based on feedback from Alejandro; add Physics and Math affiliations to header

@@
 All axioms will be physically motivated and stated as properties of ensemble spaces.
-Preliminary work shows that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can be recovered by the entropy, tightly linking the geometric and entropic structures, and that the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
+Preliminary work indicates that the symplectic form of classical mechanics and the inner product (i.e.~Born rule) of quantum mechanics can be recovered by the entropy, tightly linking the geometric and entropic structures, and that the states defined by mechanics can be understood as special cases of ensembles with the lowest possible entropy or the extreme points of the convex set.
 

### 2025-11-10 Proposals/2025-NSF-EntropicGeometry/NarrativeBody.tex
Change Project Summary based on feedback from Alejandro; add Physics and Math affiliations to header
Change file type in mention of Gabriele's biosketch as a separate document

@@
 	\textbf{An axiomatic basis for entropic structures in physical theories} \\
-	Christine A. Aidala (PI) and Alejandro Uribe (Co-PI), University of Michigan
+	Christine A. Aidala (PI, Physics) and Alejandro Uribe (Co-PI, Mathematics) \\ 
+	University of Michigan
 \end{center}
@@
 His position as U-M Physics Department staff without teaching responsibilities provides him with time and flexibility to dedicate to research related to the AoP program such as the work proposed here.
-See also his biographical sketch included as an Other Supplementary Document.\\
+See also his biographical sketch included as Other Personnel Biographical Information.\\
 

### 2025-11-13 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex

@@
 \newcommand\support{\mathrm{supp}}
-\newcommand\capacity{\mathrm{scap}}
+\newcommand\scap{\mathrm{scap}}
 \newcommand\fraction{\mathrm{frac}}
-\newcommand\size{\mathrm{frac}}
-\newcommand\frcap{\mathrm{fcap}}
 \newcommand\fcap{\mathrm{fcap}}
-\newcommand\chull{\mathrm{chull}}
 
@@
 \begin{equation}
-	\frcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\}).
+	\fcap_{\ens}(A) = \sup(\fraction_{\ens}(\hull(A))\cup\{0\}).
 \end{equation}
@@
 
-\textbf{Non-additivity of the fraction capacity.} One can then show that the fraction capacity $\frcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
+\textbf{Non-additivity of the fraction capacity.} One can then show that the fraction capacity $\fcap_{\ens} : \Sigma_{\Ens} \to [0,1]$ is a set function that satisfies the following:
 \begin{enumerate}
-	\item non-negative and unit bounded - $\frcap_{\ens}(A) \in [0,1]$
-	\item monotone - $A \subseteq B \implies \frcap_{\ens}(A) \leq \frcap_{\ens}(B)$
-	\item sub-additive - $\frcap_{\ens}(A \cup B) \leq \frcap_{\ens}(A) + \frcap_{\ens}(B)$
-	\item continuous from below and above - $\frcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \frcap_{\ens}(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
+	\item non-negative and unit bounded - $\fcap_{\ens}(A) \in [0,1]$
+	\item monotone - $A \subseteq B \implies \fcap_{\ens}(A) \leq \fcap_{\ens}(B)$
+	\item sub-additive - $\fcap_{\ens}(A \cup B) \leq \fcap_{\ens}(A) + \fcap_{\ens}(B)$
+	\item continuous from below and above - $\fcap_{\ens}(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \fcap_{\ens}(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
 \end{enumerate}
@@
 Now, the ability to express an ensemble $\ens$ as the convex combination of $M$ is precisely the ability to find a measure $p$ that represents $\ens$ such that $p(A) \neq 0$.
-That is, $\ens = \lambda \ens[a] + \bar{\lambda} \ens[b]$ with $\ens[a] \in M$ if and only if there is a probability measure $p$ that represents $\ens$ such that $p(A) = \lambda$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\frcap_{\ens}(A) = \sup(\{p(A) \, | \, p \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all measures that represent $\ens$.
+That is, $\ens = \lambda \ens[a] + \bar{\lambda} \ens[b]$ with $\ens[a] \in M$ if and only if there is a probability measure $p$ that represents $\ens$ such that $p(A) = \lambda$.\footnote{TODO: add proof somewhere} Consequently, if $M_{\ens}$ is the set of all probability measures that represent $\ens$ as defined in Choquet theory, then $\fcap_{\ens}(A) = \sup(\{p(A) \, | \, p \in M_{\ens}\})$.\footnote{TODO: prove somewhere} The fraction capacity is simply the supremum on each set over all measures that represent $\ens$.
 
@@
 We will also have $p_1(A^{\complement}) = 1- p_1(A) < 1 - p_2(A) = p_2(A^{\complement})$.
-Since the fraction capacity for a set cannot be less than the measure on that set, we have $\frcap_{\ens}(A \cup A^{\complement})=\frcap_{\ens}(X) = 1 = p_1(A) + p_1(A^{\complement}) < p_1(A) + p_2(A^{\complement}) \leq \frcap_{\ens}(A) + \frcap_{\ens}(A^{\complement})$.
+Since the fraction capacity for a set cannot be less than the measure on that set, we have $\fcap_{\ens}(A \cup A^{\complement})=\fcap_{\ens}(X) = 1 = p_1(A) + p_1(A^{\complement}) < p_1(A) + p_2(A^{\complement}) \leq \fcap_{\ens}(A) + \fcap_{\ens}(A^{\complement})$.
 That is, the fraction capacity is additive over disjoint sets of extreme points in and only in the classical case.
@@
 \begin{equation}
-\capacity(A) = \sup(2^{S(\hull(A))}\cup\{0\}).
+\scap(A) = \sup(2^{S(\hull(A))}\cup\{0\}).
 \end{equation}
@@
 
-One can then show that the state capacity $\capacity : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
+One can then show that the state capacity $\scap : \Sigma_{\Ens} \to [0,\infty]$ is a set function that satisfies the following:
 \begin{enumerate}
-	\item non-negative - $\capacity(A)\geq 0$
-	\item monotone - $A \subseteq B \implies \capacity(A) \leq \capacity(B)$
-	\item sub-additive - $\capacity(A \cup B) \leq \capacity(A) + \capacity(B)$
-	\item additive over orthogonal sets - $\capacity(A \cup B) = \capacity(A) + \capacity(B)$ if $A \ortho B$
-	\item continuous from below and above - $\capacity(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \capacity(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
+	\item non-negative - $\scap(A)\geq 0$
+	\item monotone - $A \subseteq B \implies \scap(A) \leq \scap(B)$
+	\item sub-additive - $\scap(A \cup B) \leq \scap(A) + \scap(B)$
+	\item additive over orthogonal sets - $\scap(A \cup B) = \scap(A) + \scap(B)$ if $A \ortho B$
+	\item continuous from below and above - $\scap(\lim\limits_{i \to \infty} A_i) = \lim\limits_{i \to \infty} \scap(A_i)$ for any increasing or decreasing sequence $\{A_i\}$.
 \end{enumerate}
@@
 In the discrete case, every point is associated with zero entropy, and since $2^0=1$, the state capacity reduces to the counting measure.
-In the continuous case, given a subset of phase space $A$ and $M_A$ the set of distributions with support within $A$, the maximum entropy is reached by the uniform distribution $\rho_A \in M_A$, which means we recover $\log_2 \capacity(M_A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
-Similarly, in quantum mechanics, given an orthogonal subspace $M$, the maximum entropy is recovered by the maximally mixed state, which means $\capacity(M) = 2^{\dim(M)}$.
+In the continuous case, given a subset of phase space $A$ and $M_A$ the set of distributions with support within $A$, the maximum entropy is reached by the uniform distribution $\rho_A \in M_A$, which means we recover $\log_2 \scap(M_A) = S(\rho_A)$, the fundamental postulate of statistical mechanics.
+Similarly, in quantum mechanics, given an orthogonal subspace $M$, the maximum entropy is recovered by the maximally mixed state, which means $\scap(M) = 2^{\dim(M)}$.
 
@@
 \textbf{Requirements for non-additive probability theory.} In a previous section we reviewed what types of objects we have in probability theory for classical ensembles: a sample space $X$, a probability measure $p$, random variables $F : X \to \mathbb{R}$, expectations as integrals of the probability measure $E[F] = \int_X F(x) dp$, a state counting measure $\mu$, the probability density as the Radon-Nikodym derivative $\frac{dp}{d\mu}$ and the entropy as $\int_X \rho \log \rho d\mu$.
-Given that the fraction capacity $\frcap_{\ens}$ can be seen as a generalization of the probability measure $p$ and the state capacity $\capacity$ can be seen as a generalization of the state counting measure $\mu$, is there a generalization of all other elements?
+Given that the fraction capacity $\fcap_{\ens}$ can be seen as a generalization of the probability measure $p$ and the state capacity $\scap$ can be seen as a generalization of the state counting measure $\mu$, is there a generalization of all other elements?
 That is, can we find an integral and derivative operation over non-additive measures that generalizes to all physical theories?
@@
 \begin{equation}
-	E[F] = F(\ens) = (E)\int_X F(x) d\frcap_{\ens}.
+	E[F] = F(\ens) = (E)\int_X F(x) d\fcap_{\ens}.
 \end{equation}
 which reduces to the Riemann integral in the classical (additive) case.
-The integral will depend only on the values of $F$ and $\frcap_{\ens}$ on $X$, but since these uniquely determine both $F$ and $\ens$, the integral can be understood as a map from $F$ and $\ens$ to $F(\ens)$, which is, in principle, well-defined.
+The integral will depend only on the values of $F$ and $\fcap_{\ens}$ on $X$, but since these uniquely determine both $F$ and $\ens$, the integral can be understood as a map from $F$ and $\ens$ to $F(\ens)$, which is, in principle, well-defined.
 Therefore, in principle, this integral should exist.
@@
 Additionally, we would like to be able to define an analogue of the probability density.
-We are therefore looking for a non-additive derivative $\rho = \frac{d \frcap_{\ens}}{d \capacity}$ that reduces to the Radon-Nikodym derivative in the classical (additive) case such that
-\begin{equation}
-	F(\ens) = (E)\int_X F(x) \rho(x) d\capacity
+We are therefore looking for a non-additive derivative $\rho = \frac{d \fcap_{\ens}}{d \scap}$ that reduces to the Radon-Nikodym derivative in the classical (additive) case such that
+\begin{equation}
+	F(\ens) = (E)\int_X F(x) \rho(x) d\scap
 \end{equation}
@@
 \begin{equation}
-	S(\ens) = (E)\int_X \rho(x) \log \rho(x) d\capacity.
+	S(\ens) = (E)\int_X \rho(x) \log \rho(x) d\scap.
 \end{equation}
@@
 Since $\ens[o]$ can always be obtained by the equal mixture of two opposite points (i.e.
-two orthogonal states), the fraction capacity for every singleton of an extreme point $\psi$ is $\frcap_{\ens[o]}(\{\psi\}) =\frac{1}{2}$.
+two orthogonal states), the fraction capacity for every singleton of an extreme point $\psi$ is $\fcap_{\ens[o]}(\{\psi\}) =\frac{1}{2}$.
 A statistical variable $F$ (i.e.
@@
 	\begin{aligned}
-(C)&\int F d\frcap_{\ens[o]} := \\
+(C)&\int F d\fcap_{\ens[o]} := \\
 &\int_{-\infty}^0
-(\frcap_{\ens[o]} (\{x | F (x) \geq s\})-\frcap_{\ens[o]}(X))\, ds \\
+(\fcap_{\ens[o]} (\{x | F (x) \geq s\})-\fcap_{\ens[o]}(X))\, ds \\
 &+
 \int^\infty_0
-\frcap_{\ens[o]} (\{x | F (x) \geq s\})\, ds.
+\fcap_{\ens[o]} (\{x | F (x) \geq s\})\, ds.
 	\end{aligned}
@@
 This means that $A(s)$ will include two opposite points.
-Since $\ens[o]$ can be expressed as the mixture of two opposite points, $\ens[o]$ is in the hull of $A(s)$ and therefore $\frcap_{\ens[o]}(A(s)) = 1$.
-Conversely, if $s > F(\ens[b])$, $A(s) = \emptyset$, which means $\frcap_{\ens[o]} = 0$.
+Since $\ens[o]$ can be expressed as the mixture of two opposite points, $\ens[o]$ is in the hull of $A(s)$ and therefore $\fcap_{\ens[o]}(A(s)) = 1$.
+Conversely, if $s > F(\ens[b])$, $A(s) = \emptyset$, which means $\fcap_{\ens[o]} = 0$.
 The non-trivial case, then, is when $\frac{1}{2} F(\ens[a]) + \frac{1}{2} F(\ens[b]) < s \leq F(\ens[b])$.
@@
 \begin{equation}
-	\frcap_{\ens[o]}(A(s))=
+	\fcap_{\ens[o]}(A(s))=
 	\begin{cases}
@@
 	\end{tikzpicture}
-	\caption{The integrand $\frcap_{\ens[o]}(A(s))$ as a function of $s$.}\label{fig_ChoquetIntegrand}
+	\caption{The integrand $\fcap_{\ens[o]}(A(s))$ as a function of $s$.}\label{fig_ChoquetIntegrand}
 \end{figure}
@@
 	\begin{aligned}
-		(C)&\int F d\frcap_{\ens[o]} := 
+		(C)&\int F d\fcap_{\ens[o]} := 
 		\int_{-\infty}^0
-		(\frcap_{\ens[o]} (A(s))-\frcap_{\ens[o]}(X))\, ds \\
+		(\fcap_{\ens[o]} (A(s))-\fcap_{\ens[o]}(X))\, ds \\
 		&+
 		\int^\infty_0
-		\frcap_{\ens[o]} (A(s))\, ds \\
+		\fcap_{\ens[o]} (A(s))\, ds \\
 		&= \int_{-\infty}^0 (1 - 1)ds + \int_{0}^{\frac{1}{2} a + \frac{1}{2} b}  1 ds \\
@@
 	\begin{enumerate}
-		\item non-negative: $\capacity(A) \in [0, +\infty]$
-		\item monotone: $A \subseteq B \implies \capacity(A) \leq \capacity(B)$
-		\item subadditive: $\capacity(A \cup B) \leq \capacity(A) + \capacity(B)$
-		\item additive over orthogonal sets: $A \ortho B \implies \capacity(A \cup B) = \capacity(A) + \capacity(B)$ 
+		\item non-negative: $\scap(A) \in [0, +\infty]$
+		\item monotone: $A \subseteq B \implies \scap(A) \leq \scap(B)$
+		\item subadditive: $\scap(A \cup B) \leq \scap(A) + \scap(B)$
+		\item additive over orthogonal sets: $A \ortho B \implies \scap(A \cup B) = \scap(A) + \scap(B)$ 
 	\end{enumerate}
@@
 Let $A, B \subseteq \Ens$ and let $\ens = p \ens[a] + \bar{p} \ens[b]$ for some $p \in [0,1]$, $\ens[a] \in A$ and $\ens[b] \in B$.
-By \ref{pm_es_exponentialEntropySubadditivity} and the definition of state capacity, $2^{S(\ens)} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])} \leq \capacity(A) + \capacity(B)$.
+By \ref{pm_es_exponentialEntropySubadditivity} and the definition of state capacity, $2^{S(\ens)} \leq 2^{S(\ens[a])} + 2^{S(\ens[b])} \leq \scap(A) + \scap(B)$.
 Since this is true for any finite mixture of elements of $A$ and $B$, it will be true for any element of $\hull(A \cup B)$.
 Consequently, the supremum of the exponential entropy cannot exceed the sum of the state capacities.
-Therefore $\capacity(A \cup B) \leq \capacity(A) + \capacity(B)$ (i.e.
+Therefore $\scap(A \cup B) \leq \scap(A) + \scap(B)$ (i.e.
 the state capacity is subadditive).
@@
 Let $A, B \subseteq \Ens$ be two orthogonal subsets.
-Let $\{\ens[a]_i\} \subseteq A$ be a sequence of ensembles such that $2^{S(\ens[a]_i)} \to \capacity(A)$ and let $\{\ens[b]_i\} \subseteq B$ be a sequence of ensembles such that $2^{S(\ens[b]_i)} \to \capacity(B)$.
+Let $\{\ens[a]_i\} \subseteq A$ be a sequence of ensembles such that $2^{S(\ens[a]_i)} \to \scap(A)$ and let $\{\ens[b]_i\} \subseteq B$ be a sequence of ensembles such that $2^{S(\ens[b]_i)} \to \scap(B)$.
 Consider $\ens_i = p_i \ens[a]_i + \bar{p_i} \ens[b]_i$ where $p_i = \frac{2^{S(\ens[a]_i)}}{2^{S(\ens[a]_i)} + 2^{S(\ens[b]_i)}}$.
 Then, by \ref{pm_es_exponentialEntropySubadditivity}, $2^{S(\ens_i)} = 2^{S(\ens[a]_i)} + 2^{S(\ens[b]_i)}$.
-This means that $2^{S(\ens_i)} \to \capacity(A) + \capacity(B)$.
-Therefore $\capacity(A \cup B)$ must be at least $\capacity(A) + \capacity(B)$.
-Combining with the previous result, $\capacity(A \cup B) = \capacity(A) + \capacity(B)$.
+This means that $2^{S(\ens_i)} \to \scap(A) + \scap(B)$.
+Therefore $\scap(A \cup B)$ must be at least $\scap(A) + \scap(B)$.
+Combining with the previous result, $\scap(A \cup B) = \scap(A) + \scap(B)$.
 Therefore the state capacity, as a set function, is additive over orthogonal sets of ensembles.
@@
 	Let $\{A_i\} \subseteq \Sigma_{\Ens}$ be an increasing sequence.
-Given that the state capacity is monotone, $\{\capacity(A_i)\}$ is a an increasing sequence and it will converge in the extended real line.
+Given that the state capacity is monotone, $\{\scap(A_i)\}$ is a an increasing sequence and it will converge in the extended real line.
 Let $A = \lim\limits_{i \to \infty} A_i$.
 We have $A = \cup A_i$ and $A_i \subseteq A$ for all $i$.
-Since state capacity is monotone, $\capacity(A_i) \leq \capacity(A)$ for all $i$ and therefore $\lim\limits_{i \to \infty} \capacity(A_i) \leq \capacity(A)$.
-Given any $\ens[a] \in \chull(A)$, we can find a sequence $\ens[a]_i \in \chull(A_i)$ such that $\ens[a]_i \to \ens[a]$.
+Since state capacity is monotone, $\scap(A_i) \leq \scap(A)$ for all $i$ and therefore $\lim\limits_{i \to \infty} \scap(A_i) \leq \scap(A)$.
+Given any $\ens[a] \in \overline{\hull(A)}$, we can find a sequence $\ens[a]_i \in \overline{\hull(A_i)}$ such that $\ens[a]_i \to \ens[a]$.
 Because of the continuity of the entropy, $S(\ens[a]_i) \to S(\ens[a])$.
-This means we cannot have $k$ such that $\capacity(A) > k > \capacity(A_i)$ for all $i$, and therefore $\lim\limits_{i \to \infty} \capacity(A_i) = \capacity(A)$.
+This means we cannot have $k$ such that $\scap(A) > k > \scap(A_i)$ for all $i$, and therefore $\lim\limits_{i \to \infty} \scap(A_i) = \scap(A)$.
 
@@
 Suppose $\lim\limits_{i \to \infty} \fcap_{\ens}(A_i) < \fcap_{\ens}(A)$.
-Then there would be an $\ens[a] \in \hull(A)$ such that $\size_{\ens}(\ens[a]) > \size_{\ens}(\ens[b])$ for all $\ens[b] \in \hull(A_i)$ for all $i$.
+Then there would be an $\ens[a] \in \hull(A)$ such that $\fraction_{\ens}(\ens[a]) > \fraction_{\ens}(\ens[b])$ for all $\ens[b] \in \hull(A_i)$ for all $i$.
 But if $\ens[a] \in \hull(A)$, then it can be written as a finite combination of finitely many elements of $A$.

### 2025-11-27 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
ClassicalLimit: responding to reviewers comments

@@
 \begin{equation}
-    \Delta x \Delta p \geq \Sigma(S).
+    \sigma_x \sigma_p \geq \Sigma(S).
 \end{equation}
+where $\sigma_x$ and $\sigma_p$ are the standard deviations of position and momentum and $\Sigma$ is the determinant of the covariance matrix and.
 The specific value of the uncertainty $\Sigma$ will depend on the entropy and on whether we are using classical or quantum mechanics, though the relationship will always be saturated by Gaussian states with no correlation between $x$ and $p$.
@@
 In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state~\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive.
+Moreover, even if observable do not commute in general, they commute for the maximally mixed state $\rho = I/n$ in the following sense:
+\begin{equation}
+	\<AB\>_{\rho} = \tr(AB\rho) = \tr(ABI/n) = \tr(AB)/n = \tr(BA)/n = \tr(BAI/n)= \tr(BA\rho) = \<BA\>_{\rho}
+\end{equation}
+Observables for a state close enough to the maximally mixed state, then, will ``almost commute'' in the sense that the expectation of different permutations of their product will be small.
+
 In section~\ref{sec:high_entropy_limit}, when we construct the high-entropy limit, the map that results will follow this intuition in the infinite-dimensional case, with the only difference that there is no maximally mixed state that does not change.\footnote{In the infinite-dimensional case, one could have infinite-dimensional subspaces that remain orthogonal while increasing the entropy of all its elements.
@@
 \begin{equation}
-    \frac{2 \nu^2 k_\text{B} T}{c^2} = \frac{2 \nu^2}{c^2 \beta},
+    \frac{2 \nu^2 k_rext{B} T}{c^2} = \frac{2 \nu^2}{c^2 \beta},
 \end{equation}
@@
 This explains why spin, which lives in a finite-dimensional space, is an intrinsically quantum property.
-Therefore we are going to concentrate on the case of a single degree of freedom identified by position and momentum.
-The case of multiple degrees of freedom can be recovered by increasing the entropy of all DOFs at the same time, that is, by increasing the entropy of each DOF independently, without introducing correlations.
+Therefore we are going to concentrate on the case of a single degree of freedom identified by position and momentum.\footnote{A similar limit should exist for an infinite-dimensional directional degree of freedom (i.e.
+for angular momentum).
+Though technically not a limit but more of an approximation, for finite-dimensional spaces one could look at mixed state that are close to the maximally mixed states (i.e.
+$\rho = I/n + \epsilon \sigma$).} The case of multiple degrees of freedom can be recovered by increasing the entropy of all DOFs at the same time, that is, by increasing the entropy of each DOF independently, without introducing correlations.
 
@@
 For a pure stretching map, we can provide a more specific characterization.
-Note, in fact, that a distribution with compact support is always fully characterized by all its central moments, that is, the expectations for all polynomials of position and momentum.
+First of all, given that the map will act linearly on the space of probability distributions, once we characterize its action on distributions with compact support, we fully characterize the map.
+Furthermore, a distribution with compact support is always fully characterized by all its central moments, that is, the expectations for all polynomials of position and momentum (i.e.
+the moment problem is always solvable).
 Therefore, once we know how the central moments transform through the map, we know how all distributions with compact support transform and therefore we fully characterize the map.
@@
 \end{equation}
+Given the linearity of the expectation, this relationship will remain true for all distributions that have finite moments.
+In particular, this will be true for all Schwartz functions.
+Therefore, we can use the above as the defining property of the pure stretching map.
 
@@
 where $X$ is an operator, $L_i$ are the jump operators and $\gamma_i$ are positive parameters.
-To define $T_Q$ for a single DOF we set $\gamma_i=\{\gamma\}$ and $L_i=\{a^\dagger\}$.
-We also set $H=0$ since $T_Q$ should be purely dissipative.
+To define $T_Q$ for a single DOF we set $\gamma_i=\{\gamma\}$ and $L_i=\{a^\dagger\}$.\footnote{Since $a^{\dagger} |\psi\> = a^\dagger c_n |n\> = c_n \sqrt{n+1} |n+1\>$, $|\psi\>$ is in the domain of $a^\dagger$ if and only if $\sum |c_n|^2 n < \infty$ or, equivalently, the expectation $\<\psi|n|\psi\>$ is finite.
+Since every element of the basis $|n\>$ is in the domain, $a^\dagger$ is dense in $L^2$.} We also set $H=0$ since $T_Q$ should be purely dissipative.
 The equation of motion of $a$ reduces to
@@
 axis lines=middle,
+xmin=0,
 xlabel = \(\Sigma\),
 ylabel = \(S\),
-legend style={at={(0.95,0.65)},anchor=east},
+legend style={at={(0.95,0.25)},anchor=east},
 ]
-\addplot[thick,blue,samples=150,domain=0.01:1.5] {ln(x)+1};
+\addplot[thick,blue,samples=150,domain=0.06:1.5] {ln(x)+1};
 \addlegendentry{\(\text{classical}\)}
@@
 \addlegendentry{\(\text{$\lambda = 2$}\)}
-\addplot[thick,violet,densely dashed,samples=150,domain=0.05:1.5] {((10*x+1/2)*ln(10*x+1/2)-(10*x-1/2)*ln(10*x-1/2) - ln(10)};
-\addlegendentry{\(\text{$\lambda = 10$}\)}
+\addplot[thick,violet,densely dashed,samples=150,domain=0.05:1.5] {((5*x+1/2)*ln(5*x+1/2)-(5*x-1/2)*ln(5*x-1/2) - ln(5)};
+\addlegendentry{\(\text{$\lambda = 5$}\)}
 \end{axis}
@@
     \caption{Entropy $S$ in nats for a Gaussian state as a function of uncertainty $\Sigma$, measured in units of $\hbar$.
-Between the classical and the quantum case, we have the quantum case where pure states are rescaled by a factor of $\lambda$: $S_{\lambda Q}(\Sigma) = S_Q(\lambda \Sigma) - \ln(\lambda)$.
-Taking $\lambda$ to infinity corresponds to taking $\hbar$ to zero.}
+It shows the effect of the transition from classical to quantum by shrinking the pure states by a factor of $\lambda$ instead of stretching the mixed states by the same factor.
+The rescaled uncertainty for pure states becomes $\hbar/\lambda$ and the modified Gaussian bound as becomes $S_{\lambda Q}(\Sigma) = S_Q(\lambda \Sigma) - \ln(\lambda)$.
+Taking $\lambda$ to infinity corresponds to taking $\hbar/\lambda$ to zero.}
     \label{fig:uncertainty_scaled}
@@
 If one takes the limit $\lambda \to \infty$, the above expression reduces to the classical equation.
+It should be stressed that these negative entropy states are fictitious, and it is only the difference in entropy that remains physically significant.
 
@@
 However, it will do so in the limit since $T_Q$ will become closer and closer to a pure classical stretching map.
-Lastly, we have not shown that all transformations that increase entropy must stretch phase space.
+Furthermore, we have not shown that all transformations that increase entropy must stretch phase space.
 Note, however, that a finite region of phase space can only hold mixed states with finite entropy.
 Therefore a map that does not stretch some region of phase space will necessarily lead to states whose entropy will not go to infinity as the map is reapplied over and over.
-Regardless, our ultimate goal is not to study all possible ways entropy can be increased.
+Lastly, there is the question of whether any quantum channel that increases entropy will yield the classical limit.
+If by classical limit we mean the group contraction, then the only way to achieve it is through the rescaling of the commutator between $X$ and $P$, which is exactly a stretching map as we defined it.
+More in general, however, once a state has achieved sufficiently high entropy to be approximated by a classical state, it does not matter what process was used to achieve that high-entropy state.
+The remaining difficulty will be whether the evolution can still be approximated by a Hamiltonian evolution, which is likely to fail for a generic entropy-increasing process.\footnote{In the case that a process increases entropy differently in different regions of phase space, the evolution in the limit cannot be Hamiltonian, but may still be, for example, Newtonian.} Studying this more general case is outside of the scope of this work.
 Our goal is simply to show that, as entropy increases, the classical description becomes a suitable approximation.
@@
 
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.4\textwidth]{stretchedCat.png}
+	\caption{A cat state stretched with different values of $\lambda$.
+The axes are stretched by $\sqrt{\lambda}$ as well.
+The color scale is reduced to match the range of the function, though white always corresponds to zero.
+While the positive and negative peaks have comparable magnitude initially, the stretching map dampens the negative peak much more quickly.
+Note that, consistently with Fig.
+\ref{fig:uncertainty_scaled}, the negative values are severely dampened with $\lambda=5$.}
+	\label{fig:cat_streatched}
+\end{figure}
+
 We have thus shown that, in the limit of high entropy, the Wigner distribution can be approximated by a positive function.
 We now need to show that the evolution can be approximated by classical Hamiltonian evolution.
-An intuitive way to understand why this works is to look at the evolution of the Wigner function under a Hamiltonian $H = p^2/2m + V$ where the potentials $V$ are analytic.
+An intuitive way to understand why this works is to look at the evolution of the Wigner function under a Hamiltonian $H = p^2/2m + V$ where the potential $V$ is analytic.
 We have~\cite{hillery1984distribution}
@@
 Therefore the first term becomes dominant.
+
+It should be clear that this investigation does not exhaust all possible insights and links to the now vast literature on open quantum systems, quantum statistical mechanics, quantum information, quantum sensing and so on.
+For example, entropy and entropy increase play a role in several approaches: coarse graining(\href{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.99.180403}{CITE}), decoherence(CITE), stochastic processes (CITE \href{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.129.050401}{Quantum Fokker-Planck Master Equation for Continuous Feedback Control}) and information theoretical approaches to name a few.
+On a more technical level, since both the Husimi Q and the Wigner function converge to the classical distribution, the Wehrl entropy, which is the Shannon entropy calculated on the Husimi Q, will converge to the Gibbs entropy.
+One may ask, then, whether a high Wehrl entropy is already an indicator of classicality, and, if so, it would give another insight as to why the anti-normal ordering is important.
+Lastly, the role of entropic ideas is well established in historic literature.
+It would not be possible here to comment on all of them, and the more we mention, the more the omitted mentions are felt.
+Therefore we had to only explore the most important elements to establish the limit mathematically and under the smallest conceptual footprint, also as to not distract from the main point of the paper.
 
@@
 The limit also gives us additional insights.
-As we saw, classical states for which $\Delta x \Delta p < \frac{\hbar}{e}$ are states with negative entropy, which imply a breakdown of thermodynamics.
+As we saw, classical states for which $\sigma_x \sigma_p < \frac{\hbar}{e}$ are states with negative entropy, which imply a breakdown of thermodynamics.
 It is no wonder, then, electrons cannot fall on the nucleus or that classical mechanics gives us the wrong spectra for black-body radiation.

### 2025-12-02 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
ClassicalLimit: addressing more reviewer commetns

@@
 We do not require a particular interpretation of quantum mechanics, of entropy or a particular mechanism responsible for the high entropy.
-Any will do.\footnote{The only requirement for entropy is that it matches the experimentally measured one and that its maximization corresponds to the correct experimental quantities.
+Any will do.
+\footnote{The only requirement for entropy is that it matches the experimentally measured one and that its maximization corresponds to the correct experimental quantities.
 In this vein, the Shannon/Gibbs entropy of classical mechanics and the von Neumann entropy of quantum mechanics are simply the tools we use to make predictions within the respective theories.
@@
 In the finite-dimensional case, we can have a CPTP map that increases the entropy of all states except that of the maximally mixed state, which is a fixed point.\footnote{This is a particular instance of mixing/relaxing ergodic quantum channels where the fixed point is the maximally mixed state~\cite{Burgarth2012ErgodicAM}.}  Under such a map, all pure states are mapped to mixed states, orthogonal states are mapped to non-orthogonal states and the map is strictly contractive.
-Moreover, even if observable do not commute in general, they commute for the maximally mixed state $\rho = I/n$ in the following sense:
+Moreover, even if observables do not commute in general, they commute for the maximally mixed state $\rho = I/n$ in the following sense:
 \begin{equation}
@@
 \end{equation}
-Observables for a state close enough to the maximally mixed state, then, will ``almost commute'' in the sense that the expectation of different permutations of their product will be small.
+Observables for a state close enough to the maximally mixed state, then, will ``almost commute'' in the sense that the difference in the expectation of different permutations of their product will be small.
 
@@
 In this sense the states will ``look classical.''
-
-What we described here is just a geometric property of the space of quantum mixed states, and it is independent of the source of entropy.
 
@@
 \begin{equation}
-    \frac{2 \nu^2 k_rext{B} T}{c^2} = \frac{2 \nu^2}{c^2 \beta},
+    \frac{2 \nu^2 k_\text{B} T}{c^2} = \frac{2 \nu^2}{c^2 \beta},
 \end{equation}
@@
 
+While we concentrated on the one dimensional case, the result can be generalized to multiple dimensions by performing a similar stretching operation on all degrees of freedom simultaneously.
+That is, the entropy, as it increases, it spreads equally not just within each degree of freedom, but also across degrees of freedom.
+This would give us $[T_Q(X^i),T_Q(P_i)] = \lambda \imath \hbar$, so that the same stretching factor is associated to all the appearances of $\hbar$ in the commutation relationship.
+
 To sum up, we have found a pure quantum stretching map $T_Q$.
@@
 It should be clear that this investigation does not exhaust all possible insights and links to the now vast literature on open quantum systems, quantum statistical mechanics, quantum information, quantum sensing and so on.
-For example, entropy and entropy increase play a role in several approaches: coarse graining(\href{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.99.180403}{CITE}), decoherence(CITE), stochastic processes (CITE \href{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.129.050401}{Quantum Fokker-Planck Master Equation for Continuous Feedback Control}) and information theoretical approaches to name a few.
+For example, entropy and entropy increase play a role in several approaches: coarse graining(\href{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.99.180403}{CITE}), decoherence(CITE 1 2 5 ), stochastic processes (CITE \href{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.129.050401}{Quantum Fokker-Planck Master Equation for Continuous Feedback Control}) and information-theoretic and resource-theoretic approaches (cite 3 4 6 7) to name a few.
 On a more technical level, since both the Husimi Q and the Wigner function converge to the classical distribution, the Wehrl entropy, which is the Shannon entropy calculated on the Husimi Q, will converge to the Gibbs entropy.
@@
 We have seen that classical mechanics can be recovered as the high-entropy limit of quantum mechanics.
-That is, states of higher and higher entropy are better and better approximated by classical distributions over phase space.
+That is, states of higher and higher entropy are better and better approximated by classical distributions over phase space and, if the entropy increase is done uniformly, unitary evolution can be approximated with classical Hamiltonian evolution.
 This approach to the classical limit is independent of mechanism and interpretation, as it does not matter how the entropy is increased or what one believes quantum states to represent: as long as the description is in terms of mixed states of sufficiently high entropy, classical mechanics applies.
-The approach fits naturally with experimental considerations and other approaches, such as decoherence, and recovers the established mathematical recipe, which is the group contraction for $\hbar \to 0$.
-Physically, this can be understood as taking the entropy of pure states to minus infinity, which is equivalent to saying that the relative entropy of mixed states goes to plus infinity.
+The approach fits naturally with experimental considerations and other approaches, such as decoherence, leading to an interpretation that is  thermodynamically meaningful, representation-independent, and stable under coarse-graining.
+It also recovers the established mathematical recipe, which is the group contraction for $\hbar \to 0$.\footnote{As pointed out by a reviewer, the $\hbar \to 0$ limit breaks down for chaotic or non-analytic systems, while the high entropy limit will always exist.
+Therefore there may be an additional concrete advantage to our approach over the standard one.} Physically, this can be understood as taking the entropy of pure states to minus infinity, which is equivalent to saying that the relative entropy of mixed states goes to plus infinity.
 That is, in the same way that $c \to \infty$ should be understood as $v \ll c$, $\hbar \to 0$ should be understood as $S \gg 0$, as physically we cannot take limits of physical constants.

### 2025-12-02 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
Added various references to the bib file and text

@@
 It has been shown that coherence can be maintained over long distances and among a large number of constituents, meaning that quantum systems are not necessarily small or made of few components.
-However, it is also established how coherence can be quickly lost through interaction with the environment, through decoherence~\cite{RevModPhys.75.715, SCHLOSSHAUER20191}.
+However, it is also established how coherence can be quickly lost through interaction with the environment, through decoherence~\cite{Zurek_RevModPhys.75.715, SCHLOSSHAUER20191}.
 Since decoherence increases the entropy of the system, it represents one mechanism to reach the high-entropy limit.
@@
 It should be clear that this investigation does not exhaust all possible insights and links to the now vast literature on open quantum systems, quantum statistical mechanics, quantum information, quantum sensing and so on.
-For example, entropy and entropy increase play a role in several approaches: coarse graining(\href{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.99.180403}{CITE}), decoherence(CITE 1 2 5 ), stochastic processes (CITE \href{https://journals.aps.org/prl/abstract/10.1103/PhysRevLett.129.050401}{Quantum Fokker-Planck Master Equation for Continuous Feedback Control}) and information-theoretic and resource-theoretic approaches (cite 3 4 6 7) to name a few.
+For example, entropy and entropy increase play a role in several approaches: coarse-grained measurement (e.g.~Refs.~\cite{Kofler2007,bibak2025classicallimitquantummechanics}), decoherence (e.g.~Refs.~\cite{Zurek_RevModPhys.75.715,DecoherenceBook2003,SCHLOSSHAUER20191}), stochastic processes (e.g.~Refs.~\cite{carmichael2013statistical,QuantumStochasticReview,QuantumFokkerPlanck2022}) and information-theoretic and resource-theoretic approaches (e.g.~Refs.~\cite{Goold_2016,QuantumResourceReview2019,Shiraishi2025}) to name a few.
 On a more technical level, since both the Husimi Q and the Wigner function converge to the classical distribution, the Wehrl entropy, which is the Shannon entropy calculated on the Husimi Q, will converge to the Gibbs entropy.
-One may ask, then, whether a high Wehrl entropy is already an indicator of classicality, and, if so, it would give another insight as to why the anti-normal ordering is important.
+One may ask, then, whether a high Wehrl entropy~\cite{Wehrl1979} is already an indicator of classicality, and, if so, it would give another insight as to why the anti-normal ordering is important.
 Lastly, the role of entropic ideas is well established in historic literature.
 It would not be possible here to comment on all of them, and the more we mention, the more the omitted mentions are felt.
-Therefore we had to only explore the most important elements to establish the limit mathematically and under the smallest conceptual footprint, also as to not distract from the main point of the paper.
+Therefore we only explore the most important elements to establish the limit mathematically and under the smallest conceptual footprint, also as to not distract from the main point of the paper.
 

### 2025-12-02 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
Hyphens

@@
 
-While we concentrated on the one dimensional case, the result can be generalized to multiple dimensions by performing a similar stretching operation on all degrees of freedom simultaneously.
-That is, the entropy, as it increases, it spreads equally not just within each degree of freedom, but also across degrees of freedom.
+While we concentrated on the one-dimensional case, the result can be generalized to multiple dimensions by performing a similar stretching operation on all degrees of freedom simultaneously.
+That is, the entropy, as it increases, spreads equally not just within each degree of freedom, but also across degrees of freedom.
 This would give us $[T_Q(X^i),T_Q(P_i)] = \lambda \imath \hbar$, so that the same stretching factor is associated to all the appearances of $\hbar$ in the commutation relationship.
@@
 While the positive and negative peaks have comparable magnitude initially, the stretching map dampens the negative peak much more quickly.
-Note that, consistently with Fig.
-\ref{fig:uncertainty_scaled}, the negative values are severely dampened with $\lambda=5$.}
+Note that, consistently with Fig.~\ref{fig:uncertainty_scaled}, the negative values are severely dampened with $\lambda=5$.}
 	\label{fig:cat_streatched}

### 2025-12-02 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
ClassicalLimit: added paragraph for the figure

@@
 
+In Fig.~\ref{fig:cat_streatched} we show a cat state as it is being stretched.
+Note that the stretching necessarily lowers the norm of the pseudo-probability density of both positive and negative regions.
+However, as expected, the negative regions are dampened much faster.
+After a stretching factor of $\lambda = 5$, the negative parts are negligible compared to the positive parts.
+
 We have thus shown that, in the limit of high entropy, the Wigner distribution can be approximated by a positive function.

### 2025-12-02 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
Another hyphen

@@
 The approach fits naturally with experimental considerations and other approaches, such as decoherence, leading to an interpretation that is  thermodynamically meaningful, representation-independent, and stable under coarse-graining.
-It also recovers the established mathematical recipe, which is the group contraction for $\hbar \to 0$.\footnote{As pointed out by a reviewer, the $\hbar \to 0$ limit breaks down for chaotic or non-analytic systems, while the high entropy limit will always exist.
+It also recovers the established mathematical recipe, which is the group contraction for $\hbar \to 0$.\footnote{As pointed out by a reviewer, the $\hbar \to 0$ limit breaks down for chaotic or non-analytic systems, while the high-entropy limit will always exist.
 Therefore there may be an additional concrete advantage to our approach over the standard one.} Physically, this can be understood as taking the entropy of pure states to minus infinity, which is equivalent to saying that the relative entropy of mixed states goes to plus infinity.

### 2025-12-02 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
ClassicalLimit: fixed figure for landscape

@@
 \end{equation}
-where $\sigma_x$ and $\sigma_p$ are the standard deviations of position and momentum and $\Sigma$ is the determinant of the covariance matrix and.
+where $\sigma_x$ and $\sigma_p$ are the standard deviations of position and momentum and $\Sigma$ is the determinant of the covariance matrix.
 The specific value of the uncertainty $\Sigma$ will depend on the entropy and on whether we are using classical or quantum mechanics, though the relationship will always be saturated by Gaussian states with no correlation between $x$ and $p$.
@@
 for angular momentum).
-Though technically not a limit but more of an approximation, for finite-dimensional spaces one could look at mixed state that are close to the maximally mixed states (i.e.
+Though technically not a limit but more of an approximation, for finite-dimensional spaces one could look at mixed states that are close to the maximally mixed states (i.e.
 $\rho = I/n + \epsilon \sigma$).} The case of multiple degrees of freedom can be recovered by increasing the entropy of all DOFs at the same time, that is, by increasing the entropy of each DOF independently, without introducing correlations.
@@
 It shows the effect of the transition from classical to quantum by shrinking the pure states by a factor of $\lambda$ instead of stretching the mixed states by the same factor.
-The rescaled uncertainty for pure states becomes $\hbar/\lambda$ and the modified Gaussian bound as becomes $S_{\lambda Q}(\Sigma) = S_Q(\lambda \Sigma) - \ln(\lambda)$.
+The rescaled uncertainty for pure states becomes $\hbar/\lambda$ and the modified Gaussian bound becomes $S_{\lambda Q}(\Sigma) = S_Q(\lambda \Sigma) - \ln(\lambda)$.
 Taking $\lambda$ to infinity corresponds to taking $\hbar/\lambda$ to zero.}
@@
 	\centering
-	\includegraphics[width=0.4\textwidth]{stretchedCat.png}
+	\includegraphics[width=\textwidth]{stretchedCat.png}
 	\caption{A cat state stretched with different values of $\lambda$.
@@
 The approach fits naturally with experimental considerations and other approaches, such as decoherence, leading to an interpretation that is  thermodynamically meaningful, representation-independent, and stable under coarse-graining.
-It also recovers the established mathematical recipe, which is the group contraction for $\hbar \to 0$.\footnote{As pointed out by a reviewer, the $\hbar \to 0$ limit breaks down for chaotic or non-analytic systems, while the high-entropy limit will always exist.
+It also recovers the established mathematical recipe, which is the group contraction for $\hbar \to 0$.\footnote{As pointed out by a reviewer, the $\hbar \to 0$ limit can be problematic for chaotic or non-analytic systems.
 Therefore there may be an additional concrete advantage to our approach over the standard one.} Physically, this can be understood as taking the entropy of pure states to minus infinity, which is equivalent to saying that the relative entropy of mixed states goes to plus infinity.

### 2025-12-06 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
Change AoP from a project to a program

@@
 \section*{Acknowledgements}
-This paper is part of the ongoing Assumptions of Physics project~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
+This paper is part of the research program Assumptions of Physics~\cite{aop-book}, which aims to identify a handful of physical principles from which the basic laws can be rigorously derived.
 
-% TODO: include funding information
 \paragraph{Funding information}

### 2025-12-06 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
Delete extra space before footnote

@@
 We do not require a particular interpretation of quantum mechanics, of entropy or a particular mechanism responsible for the high entropy.
-Any will do.
-\footnote{The only requirement for entropy is that it matches the experimentally measured one and that its maximization corresponds to the correct experimental quantities.
+Any will do.\footnote{The only requirement for entropy is that it matches the experimentally measured one and that its maximization corresponds to the correct experimental quantities.
 In this vein, the Shannon/Gibbs entropy of classical mechanics and the von Neumann entropy of quantum mechanics are simply the tools we use to make predictions within the respective theories.

### 2025-12-09 Articles/2024-ClassicalLimit/classicalLimitArticle.tex
Clean copy for Physica Scripta, with corresponding author
Added hyphens, ~ after i.e., moved Wehrl entropy citation one sentence earlier, fixed refs (added preprint for one not yet published, removed months)

@@
 \author[2]{Manuele Landini}
-\author[1]{Christine A. Aidala}
+\author[1]{Christine A. Aidala\thanks{Corresponding author.~\href{mailto:caidala@umich.edu}{caidala@umich.edu}}}
 
@@
 In systems with finitely many possible states we can find the stationary point of $S$ as a function of $E$ for which $1/T=\partial S/\partial E$ is zero.
-This stationary point represents the maximum entropy reachable by the system, and therefore there is no high entropy limit.
+This stationary point represents the maximum entropy reachable by the system, and therefore there is no high-entropy limit.
 This is exactly the case where temperature can have negative values.
@@
 This explains why spin, which lives in a finite-dimensional space, is an intrinsically quantum property.
-Therefore we are going to concentrate on the case of a single degree of freedom identified by position and momentum.\footnote{A similar limit should exist for an infinite-dimensional directional degree of freedom (i.e.
-for angular momentum).
-Though technically not a limit but more of an approximation, for finite-dimensional spaces one could look at mixed states that are close to the maximally mixed states (i.e.
-$\rho = I/n + \epsilon \sigma$).} The case of multiple degrees of freedom can be recovered by increasing the entropy of all DOFs at the same time, that is, by increasing the entropy of each DOF independently, without introducing correlations.
+Therefore we are going to concentrate on the case of a single degree of freedom identified by position and momentum.\footnote{A similar limit should exist for an infinite-dimensional directional degree of freedom (i.e.~for angular momentum).
+Though technically not a limit but more of an approximation, for finite-dimensional spaces one could look at mixed states that are close to the maximally mixed states (i.e.~$\rho = I/n + \epsilon \sigma$).} The case of multiple degrees of freedom can be recovered by increasing the entropy of all DOFs at the same time, that is, by increasing the entropy of each DOF independently, without introducing correlations.
 
@@
     \item Show that it can be defined over anti-normal ordering.
-Show that if we rescale the zero of entropy, the commutators become $[X, P]= \frac{\imath \hbar}{\lambda}$, which makes the limit $\lambda \to \infty$ mathematically equivalent to $\hbar \to 0$
+Show that if we rescale the zero of entropy, the commutators become $[X, P]= \frac{\imath \hbar}{\lambda}$, which makes the limit $\lambda \to \infty$ mathematically equivalent to $\hbar \to 0$.
     \item Show how, in the limit, the Wigner $W$ and Husimi $Q$ distributions become closer and closer.
@@
 First of all, given that the map will act linearly on the space of probability distributions, once we characterize its action on distributions with compact support, we fully characterize the map.
-Furthermore, a distribution with compact support is always fully characterized by all its central moments, that is, the expectations for all polynomials of position and momentum (i.e.
-the moment problem is always solvable).
+Furthermore, a distribution with compact support is always fully characterized by all its central moments, that is, the expectations for all polynomials of position and momentum (i.e.~the moment problem is always solvable).
 Therefore, once we know how the central moments transform through the map, we know how all distributions with compact support transform and therefore we fully characterize the map.
@@
 Therefore, while negative regions can be in principle found at any finite $\lambda$, $W$ tends to a positive function in the limit.
-As usual for the $W$ distribution, the phase space size of negative regions is limited to $\hbar$ by the uncertainty principle.
+As usual for the $W$ distribution, the phase-space size of negative regions is limited to $\hbar$ by the uncertainty principle.
 The weight of the function in such regions is limited between $\pm 2/\hbar$ for pure states.
@@
 For example, entropy and entropy increase play a role in several approaches: coarse-grained measurement (e.g.~Refs.~\cite{Kofler2007,bibak2025classicallimitquantummechanics}), decoherence (e.g.~Refs.~\cite{Zurek_RevModPhys.75.715,DecoherenceBook2003,SCHLOSSHAUER20191}), stochastic processes (e.g.~Refs.~\cite{carmichael2013statistical,QuantumStochasticReview,QuantumFokkerPlanck2022}) and information-theoretic and resource-theoretic approaches (e.g.~Refs.~\cite{Goold_2016,QuantumResourceReview2019,Shiraishi2025}) to name a few.
-On a more technical level, since both the Husimi Q and the Wigner function converge to the classical distribution, the Wehrl entropy, which is the Shannon entropy calculated on the Husimi Q, will converge to the Gibbs entropy.
-One may ask, then, whether a high Wehrl entropy~\cite{Wehrl1979} is already an indicator of classicality, and, if so, it would give another insight as to why the anti-normal ordering is important.
+On a more technical level, since both the Husimi Q and the Wigner function converge to the classical distribution, the Wehrl entropy~\cite{Wehrl1979}, which is the Shannon entropy calculated on the Husimi Q, will converge to the Gibbs entropy.
+One may ask, then, whether a high Wehrl entropy is already an indicator of classicality, and, if so, it would give another insight as to why the anti-normal ordering is important.
 Lastly, the role of entropic ideas is well established in historic literature.
@@
 Classical mechanics fails at low entropy.
-The use of the correspondence principle as guidance for the development of quantum mechanics, then, can be understood as requiring that the new quantized theory reproduces classical mechanics at high entropy.
+The use of the correspondence principle as guidance for the development of quantum mechanics, then, can be understood as requiring that the new quantized theory reproduce classical mechanics at high entropy.
 

### 2025-12-09 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewed introduction

--- 2025-08-29 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
@@
 
-\title{A non-additive generalization of probability theory \\for quantum mechanics and beyond}
+\title{A non-additive generalization of measure theory and \\ probability theory for quantum mechanics and beyond}
 \author{Gabriele Carcassi}
@@
 In the past few decades, there has been a growing interest in developing approaches that generalize the notion of states and physical theories in terms of probabilistic and/or entropic concepts \cite{plavala2023,wilce2025,ariano2017,caticha2019,hall2016}, which stems from the underlying aim to characterize physical theories from an operational viewpoint.
-One basic insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
+A key insight is that statistical/probabilistic concepts often provide a suitable basis for this generalization, and it is in this setting where most of these approaches operate.
+
 A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969,fine1982}.
-Quasiprobabilistic approaches \cite{zachos2005,mandel1995,monchietti2025} are also used in quantum mechanics, but it is not completely clear how exactly they relate to the standard one.
-
-The paper aims to ask the following question: can we, by arguing from necessity, find a minimal set of physical requirements for a state space of a physical system that allows us to directly connect to all the measure theoretic structures currently used in physics (i.e.
-probability measures, Liouville measure to count states, quasiprobability, ...) and point to a straightforward generalization of such structures?
-We will show that this is indeed possible.\footnote{Note that we believe that the requirements we set, though necessary, are not sufficient to generalize the Lie algebra given by the classical Poisson brackets and the quantum commutators, meaning they cannot fully recover mechanics.
+One possible generalization is through quasiprobability,\cite{zachos2005,mandel1995,monchietti2025} which signed probability measures.
+However, given that one can non-negative quasiprobability representations of quantum spaces, the line between standard probability and quasiprobability has not been adequately established.
+Moreover, the measure that quantities the number of states in a region is also fundamental, as it is used both to define probability density and to define the notion of the entropy.
+Currently, a quantum analog of the count of states has not been properly identified.
+
+In this work we aim to fully explore the connection between physical theories and measure theoretic tools.
+Arguing from necessity, we want to find a minimal set of physical requirements for a state space of a physical system that allows us to directly connect to all the measure theoretic structures currently used in physics (i.e.
+probability measures, Liouville measure to count states, quasiprobability, ...) and point to a straightforward generalization of such structures.\footnote{Note that we believe that the requirements we set, though necessary, are not sufficient to generalize the Lie algebra given by the classical Poisson brackets and the quantum commutators, meaning they cannot fully recover mechanics.
 Nonetheless, they are enough to recover measure theoretic structures, the topic of this paper.}
@@
 there is a family of affine functions that identify each ensemble through their values).
-Mathematically, these premises lead us to assume that an ensemble space is a convex subset of a metrizable locally convex topological vector space.
+Mathematically, these premises lead us to assume that an ensemble space is a convex subset of a metrizable second countable locally convex topological vector space.
 The question is, then, what generalized measure theoretic tool can quantify the number of states (reducing to the Liouville measure in the classical case) or the number of possible outcomes of a measurement (reducing to the dimension of a subspace in the quantum case)?
@@
 The above two approaches exhaust what can be done with additive measures, so we turn to the non-additive case.
-Given an ensemble $\ens$, we can ask what fraction can be produced by a mixture of a given set of ensembles.
-This leads to the fraction capacity: a sub-additive set function that, like the probability measure, is bounded by one, monotonic and continuous from above and below.
+We found a general construction that, given a function $f: \Ens \to \mathbb{R}$, provides us with a set function $cs_f(A)$ that returns the supremum of $f$ over the convex hull of $A$.
+This convex supremum is always a monotone set function continuous from below.
+The generalization of both probability measures and state counting measures are applications of this construction.
+
+To generalize the notion of probability measure, given a target ensemble $\ens$, we can define the fraction $\fraction_{\ens}(\ens[a])$ to be the highest possible $p$ such that $\ens = p \ens[a] + (1-p) \ens[b]$.
+The fraction capacity $\frcap_{\ens}(A)$ is the corresponding convex supremum, which returns what fraction of $\ens$ can be produced by a mixture of a given set of ensembles $A$.
+The fraction capacity is found to be a monotone, non-negative, bounded by one, continuous from below, sub-additive set function.
 Over the extreme points, the fraction capacity represents the supremum of all possible measures that can represent the same ensemble in Choquet theory.
@@
 
-We can also generalize a notion of count of states.
-While in classical statistical mechanics we start with a measure that quantifies the count of states and define the entropy using its logarithm, for the ensemble space we proceed in the opposite way: we start from the entropy and we define the count of states as the exponential of the entropy (i.e.
-the entropic volume)\cite{hall1999}.
-From this, we can define a state capacity, which is again a sub-additive measure.
-It reduces to the Liouville measure in the classical case and the dimensionality of subspaces in the quantum case.
+To generalize the notion of count of states, we note that in classical statistical mechanics the entropy over uniform distribution is the logarithm of the count of states in the region, which is the supremum of the entropy of all measures with the same support.
+We proceed in the opposite way: for each ensemble we can define the exponential of the entropy (i.e.
+its entropic volume)\cite{hall1999} and define the state capacity to be the corresponding convex supremum.
+The state capacity is found to be monotone, non-negative, continuous from below, sub-additive set function.
+It reduces to the Liouville measure in the classical case and to the dimensionality of subspaces in the quantum case.
 

### 2025-12-09 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: aligning text with more recent ideas

@@
 	%Given a generic space of ensembles, we can define the fraction capacity as the maximum fraction of a particular ensemble that can be understood as a mixture of ensembles from a given set.
-This gives a non-additive (i.e.
-fuzzy) measure that reduces to a probability (i.e.
-additive) measure for classical spaces and for quantum measurements.
+This gives a non-additive (i.e.~fuzzy) measure that reduces to a probability (i.e.~additive) measure for classical spaces and for quantum measurements.
 We can also define the state capacity as the exponential of the maximum entropy reachable by a mixture of ensembles from a given set.
@@
 
-In this work we aim to fully explore the connection between physical theories and measure theoretic tools.
-Arguing from necessity, we want to find a minimal set of physical requirements for a state space of a physical system that allows us to directly connect to all the measure theoretic structures currently used in physics (i.e.
-probability measures, Liouville measure to count states, quasiprobability, ...) and point to a straightforward generalization of such structures.\footnote{Note that we believe that the requirements we set, though necessary, are not sufficient to generalize the Lie algebra given by the classical Poisson brackets and the quantum commutators, meaning they cannot fully recover mechanics.
+\textbf{Article aim.} In this work we aim to fully explore the connection between physical theories and measure theoretic tools.
+Arguing from necessity, we want to find a minimal set of physical requirements for a state space of a physical system that allows us to directly connect to all the measure theoretic structures currently used in physics (i.e.~probability measures, Liouville measure to count states, quasiprobability, ...) and point to a straightforward generalization of such structures.\footnote{Note that we believe that the requirements we set, though necessary, are not sufficient to generalize the Lie algebra given by the classical Poisson brackets and the quantum commutators, meaning they cannot fully recover mechanics.
 Nonetheless, they are enough to recover measure theoretic structures, the topic of this paper.}
@@
 The starting point is that any physical law deals with experimentally reproducible relationships.
-This implies a physical theory must give us the space of statistical ensembles (i.e.
-the collection of all possible preparation procedures) that the theory allows.
+This implies a physical theory must give us the space of statistical ensembles (i.e.~the collection of all possible preparation procedures) that the theory allows.
 As a basic requirement, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from Generalized Probabilistic Theories (GPTs) is the entropy requirement.
-We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough experimentally measurable quantities to identify each ensemble (i.e.
-there is a family of affine functions that identify each ensemble through their values).
+We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough experimentally measurable quantities to identify each ensemble (i.e.~there is a family of affine functions that identify each ensemble through their values).
 Mathematically, these premises lead us to assume that an ensemble space is a convex subset of a metrizable second countable locally convex topological vector space.
@@
 
+\textbf{Main finding.}  Choquet theory gives us a representation of ensembles in terms of convex combination of extreme points (i.e.~pure states).
+Classical spaces are those for which this representation is unique.
+Psuedo-probability measures are affine representations of the ensemble space, and different representation chosen based on the problem at hand.
+The probability measure finds a natural non-additive generalization in the fraction capacity $\frcap_{\ens}(A)$, which returns the highest possible fraction that an ensemble $\ens$ can be recovered from a convex combination of elements of $A$.
+The state counting measures finds a natural non-additive generalization in the state capacity $\capacity(A)$ which returns the highest exponential of the entropy (i.e.~entropic volume) achievable with a convex combination of elements of $A$.
+
+\section{Article outline}
+
 We will first review how measure theory is used in classical probability \cite{kolmogorov2018} and classical statistical mechanics \cite{Peliti2011}.
-We will connect this to Choquet theory \cite{phelps2001}, which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.
-each ensemble can be represented by a probability measure over pure states).
+We will connect this to Choquet theory \cite{phelps2001}, which aims to represent each point of a convex subset as a probability measure over its extreme points (i.e.~each ensemble can be represented by a probability measure over pure states).
 While such representation is always possible, at least in the compact case, it is not unique.
@@
 To generalize the notion of count of states, we note that in classical statistical mechanics the entropy over uniform distribution is the logarithm of the count of states in the region, which is the supremum of the entropy of all measures with the same support.
-We proceed in the opposite way: for each ensemble we can define the exponential of the entropy (i.e.
-its entropic volume)\cite{hall1999} and define the state capacity to be the corresponding convex supremum.
+We proceed in the opposite way: for each ensemble we can define the exponential of the entropy (i.e.~its entropic volume)\cite{hall1999} and define the state capacity to be the corresponding convex supremum.
 The state capacity is found to be monotone, non-negative, continuous from below, sub-additive set function.
@@
 \end{equation}
-where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{The Shannon entropy is recovered as the only upper bound possible.}  Intuitively, the variability (i.e.
-the entropy) cannot decrease during mixing, and will have an upper bound when mixing two ensembles whose instances are always distinguishable.
+where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{The Shannon entropy is recovered as the only upper bound possible.}  Intuitively, the variability (i.e.~the entropy) cannot decrease during mixing, and will have an upper bound when mixing two ensembles whose instances are always distinguishable.
 
@@
 No ensemble is separate from itself.
-An internal point is an ensemble that is not separate from any ensemble (i.e.
-it is the mixture of any ensemble).
-An extreme point is an ensemble that is separate from all others (i.e.
-it is only a mixture of itself).
+An internal point is an ensemble that is not separate from any ensemble (i.e.~it is the mixture of any ensemble).
+An extreme point is an ensemble that is separate from all others (i.e.~it is only a mixture of itself).
 
@@
 In quantum mechanics, these two notions are different: any two pure states are separate but are not necessarily orthogonal.
-For example, pure ensembles of spin up and spin down are orthogonal (i.e.
-mutually exclusive) precisely because it takes one instance to determine which one is which.
+For example, pure ensembles of spin up and spin down are orthogonal (i.e.~mutually exclusive) precisely because it takes one instance to determine which one is which.
 Pure ensembles of spin up and spin right, however, are not orthogonal precisely because their instances cannot be perfectly discriminated.
@@
 
-\textbf{Classical probability measures.} In the classical case, given the topological space of possible states $X$ (i.e.
-the sample space), ensembles are defined as probability measures $p : \Sigma_X \to [0,1]$.
+\textbf{Classical probability measures.} In the classical case, given the topological space of possible states $X$ (i.e.~the sample space), ensembles are defined as probability measures $p : \Sigma_X \to [0,1]$.
 That is, a probability measure assigns to each element of the Borel algebra $\Sigma_X$\footnote{In our overall project, Borel sets represent statements associated with a test, regardless of termination.}, an event, the probability to find it to be true.
@@
 That is, we are able to recognize whether a particular die landed or didn't land on a particular face.} and, consequently, the power set is the $\sigma$-algebra.
-In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold.\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variables which are verifiable up to finite but arbitrarily small precision.}  The probability measure is countably additive, meaning that, given a countable collection $A_i$ of pair-wise disjoint events (i.e.
-$A_i \cap A_j = \emptyset$), the probability of the union is the sum of the probability of the respective events:
+In the continuous case, $X$ will be classical phase space endowed with the usual topology of a manifold.\footnote{Physically, the topology of a manifold corresponds to the ability to label each possible case through a set of continuous variables which are verifiable up to finite but arbitrarily small precision.}  The probability measure is countably additive, meaning that, given a countable collection $A_i$ of pair-wise disjoint events (i.e.~$A_i \cap A_j = \emptyset$), the probability of the union is the sum of the probability of the respective events:
 \begin{equation}
@@
 \end{equation}
-The expectation is linear with respect to the variable (i.e.
-$E[aF+bG]= aE[F]+bE[G]$).
+The expectation is linear with respect to the variable (i.e.~$E[aF+bG]= aE[F]+bE[G]$).
 In our case, what is more interesting is that it is an affine of the probability measure: if $p = \lambda p_1 + \bar{\lambda} p_2$ then $\int_X F dp = \lambda \int_X F dp_1 + \bar{\lambda} \int_X F dp_2$.
@@
 Let $\mu : \Sigma_X \to [0, + \infty]$ be a measure on $X$ that quantifies the count of states in each region.
-In the discrete case this is simply the counting measure (i.e.
-three possible faces of the die correspond to three possible states), while in the continuous case this corresponds to the Liouville measure (i.e.
-volumes in phase space correspond to the count of states).
+In the discrete case this is simply the counting measure (i.e.~three possible faces of the die correspond to three possible states), while in the continuous case this corresponds to the Liouville measure (i.e.~volumes in phase space correspond to the count of states).
 This measure is additive over disjoint regions as they represent mutually exclusive events.
@@
 This is a mischaracterization.\cite{ferrie2010,adhikary2020} For example, the Husimi Q representation provides us with a non-negative quasiprobability function.
-That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.\footnote{Since the Husimi Q is defined as $Q_{\rho} ( \alpha ) = \frac{1}{\pi} \langle \alpha | \rho | \alpha \rangle$, it is proportional to the probability of measuring a coherent (i.e.
-Gaussian) state $| \alpha \rangle$, which is non-negative.
+That is, the Husimi Q is a non-negative density distribution that integrates to one, yet it is not a classical probability distribution.\footnote{Since the Husimi Q is defined as $Q_{\rho} ( \alpha ) = \frac{1}{\pi} \langle \alpha | \rho | \alpha \rangle$, it is proportional to the probability of measuring a coherent (i.e.~Gaussian) state $| \alpha \rangle$, which is non-negative.
 } Similarly, a SIC-POVM characterizes the Bloch ball by four pure states that form a tetrahedron inscribed in the ball.
@@
 Instead, the crucial difference is that in classical probability all points of the sample space, and therefore disjoint regions, represent mutually exclusive events, and all mutually exclusive events are represented by disjoint regions.
-Mutually exclusive events can be prepared and observed independently from the others (i.e.
-they are orthogonal in the sense that they maximize entropy increase during mixture); we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasiprobability in quantum mechanics.
+Mutually exclusive events can be prepared and observed independently from the others (i.e.~they are orthogonal in the sense that they maximize entropy increase during mixture); we can assign some probability to some region without having to assign it to another disjoint region.\footnote{The only constraint is that the probability over the whole space is one.} This is not the case for quasiprobability in quantum mechanics.
 
@@
 The only way for $A$ and $B$ to be mutually exclusive, then, is if different measures represent different ensembles, which means the ensemble space is a simplex.
-This tells us that we can use classical probability exactly when all extreme points are mutually exclusive (i.e.
-orthogonal as defined from the entropy) and the ensemble space is a simplex.
+This tells us that we can use classical probability exactly when all extreme points are mutually exclusive (i.e.~orthogonal as defined from the entropy) and the ensemble space is a simplex.
 
@@
 
-% 4 pros and cons - can choose a repretation for the problem at hand; no fundamental representation; boundaries ore not well defined (i.e. which functions represnt valid states)
+% 4 pros and cons - can choose a repretation for the problem at hand; no fundamental representation; boundaries ore not well defined (i.e.~which functions represnt valid states)
 
@@
 In this case, the quasiprobability associated to each ensemble is unique, but the space over which the quasiprobability is defined is not.
+
+\section{Convex supremum}
+
+The set functions we are going to construct both can be understood as coming from the same construction that has very general applicability.
+Let $\Ens$ be a convex space and $f : \Ens \to \mathbb{R}$ a real-valued function defined on it.
+Recall that, given a subset $A \subseteq \Ens$, its convex hull $\hull(A)$ includes all possible convex combinations of elements of $A$.
+Given a set $A$, then, we can ask what is the highest value that $f$ can reach over $\hull(A)$.
+If $A$ is the empty set, this would not be defined, so we need a value $f_{\emptyset}$ for the empty set such that $f_{\emptyset} \leq \inf(f(\Ens))$ to preserve monotonicity.
+Given $f$ and $f_{\emptyset}$, we can define the convex supreme as $cs_f(A) = \sup(f(\hull(A)) \cup \{f_{\emptyset}\})$.
+This will be a monotone set function continuous from below and with a range of $[f_{\emptyset}, \sup(f(\Ens))]$.
+
+To put these properties in context, a monotone set function continuous from below is a measure if and only if it is additionally non-negative and finitely additive.
+Therefore, this construction naturally gives well behaved non-additive measures and, since our measures come from this construction, it may be useful to study this class of set functions more in general.
 
@@
 However, the question of what is the biggest mixture coefficient one can use in a specific case is well-defined, and generalizes to the non-compact case as well.
-Therefore, our generalization of probability is not in terms of what an ensemble can lead to (i.e.
-measurements), but in terms of how ensembles can be produced (i.e.
-preparations).
+Therefore, our generalization of probability is not in terms of what an ensemble can lead to (i.e.~measurements), but in terms of how ensembles can be produced (i.e.~preparations).
 
@@
 We now extend this idea from a single ensemble $\ens[a] \in \Ens$ to a Borel set of ensembles $A \subseteq \Ens$.
-Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.
-mixtures), we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
+Given that the $\hull(A)$ is the set of all possible convex combinations (i.e.~mixtures), we define the \textbf{fraction capacity} of $A$ for $\ens$ to be
 \begin{equation}
@@
 \end{equation}
-This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$.
+This returns the biggest fraction of $\ens$ that can be achieved with a mixture of elements of $A$. This is exactly the convex supremum of the fraction with $f_{\emptyset} = 0$.
 
@@
 It is therefore a non-additive probability measure.
-The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.
-fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\cite{nonadditive2014}
+The name fraction capacity is chosen to both signify the ability of the set $A$ to contain $\ens$ (i.e.~fraction capacity of one means $\ens$ is within the convex combinations of $A$) and to indicate that it will be a non-additive monotonic measure, which are called ``capacities'' in some literature.\cite{nonadditive2014}
 
@@
 On the other hand, unlike Choquet theory and like quasiprobability, once the space is fixed, there is one unique fraction capacity for each ensemble.
-Moreover, the setup is more general than Choquet theory (i.e.
-the ensemble space need not be compact) and stems from necessary physical requirements.
+Moreover, the setup is more general than Choquet theory (i.e.~the ensemble space need not be compact) and stems from necessary physical requirements.
 An in-depth study on fraction capacities, then, is guaranteed to be applicable to all physical theories, including those yet to be formulated.
@@
 This returns the upper bound to the exponential of entropy achievable with a mixture of elements of $A$.
+This is exactly the convex supremum of the exponential of the entropy with $f_{\emptyset} = 0$.
 
@@
 
+\section{Quantum measures and quantum logic}
+
+Note that our approach also connects, in a sense, to what is done in quantum measure theory and quantum logic.(TODO cite) In these approaches, the probability measure is not defined on Borel set, but rather on the lattice of closed subspaces.
+While we work on the ensemble space rather than space of pure states, the notion of subspace in two spaces is equivalent.
+If we restrict the state capacity on the lattice of orthogonal subspaces, then, one has $\capacity(X \vee Y) = \capacity(X) + \capacity (Y)$ if $X \perp Y$.
+That is, the measure is additive over orthogonal sets, instead of disjoint, sets.
+Similarly, given an observable $O$, one can define the set $\mathcal{C}$ of all subspaces that correspond to the Borel sets of $O$.
+In the ensemble space, these correspond to the closed convex subspaces of all possible mixtures of elements whose wavefunction have support within the corresponding Borel set.
+All $X,Y \in \mathcal{C}$ will either orthogonal (i.e.
+$X \perp Y$) or one a subspace of the other (i.e.
+$X \subseteq Y$ or $Y \subseteq X$).
+If a density operator commutes of $O$, then we can always express it as a mixture of elements of $X$ and $X^{\perp}$ for all $X \in \mathcal{C}$.
+In this case, $\frcap_{\ens}$ is additive if restricted on $\mathcal{C}$, which means we recover the probability measures associated with the observable.
+
+It is likely that there are further connection that could be made, but since quantum measures and quantum logic are not, strictly speaking, measure theoretic tools, we will leave this for future work.
+
 \section{Non-additive generalization of probability theory}
@@
 
-The Sugeno integral does not recover the Lebesgue integral in the case the capacity is additive (i.e.
-classical), and is therefore quickly ruled out.
+The Sugeno integral does not recover the Lebesgue integral in the case the capacity is additive (i.e.~classical), and is therefore quickly ruled out.
 The Choquet integral does, so it will work in the classical case.
@@
 The center point $\ens[o]$ is the maximally mixed state.
-Since $\ens[o]$ can always be obtained by the equal mixture of two opposite points (i.e.
-two orthogonal states), the fraction capacity for every singleton of an extreme point $\psi$ is $\frcap_{\ens[o]}(\{\psi\}) =\frac{1}{2}$.
-A statistical variable $F$ (i.e.
-an observable), is an affine function of the ensemble space.
-This means that two opposite points on the surface $\ens[a]$ and $\ens[b]$ (i.e.
-the eigenstates) will take the minimum and maximum values $a = F(\ens[a])$ and $b = F(\ens[b])$.
+Since $\ens[o]$ can always be obtained by the equal mixture of two opposite points (i.e.~two orthogonal states), the fraction capacity for every singleton of an extreme point $\psi$ is $\frcap_{\ens[o]}(\{\psi\}) =\frac{1}{2}$.
+A statistical variable $F$ (i.e.~an observable), is an affine function of the ensemble space.
+This means that two opposite points on the surface $\ens[a]$ and $\ens[b]$ (i.e.~the eigenstates) will take the minimum and maximum values $a = F(\ens[a])$ and $b = F(\ens[b])$.
 For simplicity, we can picture $\ens[a]$ and $\ens[b]$ as the extreme points of the vertical axis and we will assume $0<a<b$.
@@
 Conversely, all statistical variables are linear on the ensemble space, so the restriction on the extreme points cannot be arbitrary.
-In the Bloch ball case, for example, all linear functions are fully determined by two opposite points on the sphere and by two values (i.e.
-the eigenstates and the eigenvalues of the corresponding operator).
+In the Bloch ball case, for example, all linear functions are fully determined by two opposite points on the sphere and by two values (i.e.~the eigenstates and the eigenvalues of the corresponding operator).
 The value for all the other points is determined.
@@
 In principle, these non-additive measures should allow a generalization of probability calculus that reduces to classical probability in the classical case.
-We have seen, however, that the two standard non-additive integrals (i.e.
-the Sugeno and the Choquet integral) are not suitable.
+We have seen, however, that the two standard non-additive integrals (i.e.~the Sugeno and the Choquet integral) are not suitable.
 It will likely need a broader collaboration to build these tools.

### 2025-12-17 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: reviewing draft

@@
 
-\title{A non-additive generalization of measure theory and \\ probability theory for quantum mechanics and beyond}
+\title{A non-additive generalization of probability and \\ state count for all physical theories}
 \author{Gabriele Carcassi}
@@
 \begin{abstract}
-	We present a physically motivated generalization of probability theory that is suitable for classical mechanics, quantum mechanics and any future physical theory that allows a statistical description.
-The goal is to put the use of classical and quantum probability in a broader context, and show how the current mathematical structures are likely not suitable to solve the open problems in the foundations of physics.
-For the more mathematically inclined, we will point to areas where new math or generalization of established math are needed.
-For the more philosophically incline, we will point to areas where further conceptual work is needed.
+	We present a physically motivated non-additive generalization of probability and state counting measures that is equally valid on all physical theories that allow a statistical description.
+In the classical case, these reduce to the standard probability and Liouville/counting measures, but retain useful properties in the general case.
+Both constructions are instances of a more general construction that may may be of more general interest.
+Moreover, the expectation for statistical variables is not recovered by standard non-additive integrals, and therefore new techniques will need to be developed.
 
@@
 A key problem, then, is what notion of probability should be used given that the standard measure theoretic Kolmogorovian approach does not generally work in quantum mechanics\cite{chsh1969,fine1982}.
-One possible generalization is through quasiprobability,\cite{zachos2005,mandel1995,monchietti2025} which signed probability measures.
-However, given that one can non-negative quasiprobability representations of quantum spaces, the line between standard probability and quasiprobability has not been adequately established.
-Moreover, the measure that quantities the number of states in a region is also fundamental, as it is used both to define probability density and to define the notion of the entropy.
-Currently, a quantum analog of the count of states has not been properly identified.
-
-\textbf{Article aim.} In this work we aim to fully explore the connection between physical theories and measure theoretic tools.
-Arguing from necessity, we want to find a minimal set of physical requirements for a state space of a physical system that allows us to directly connect to all the measure theoretic structures currently used in physics (i.e.~probability measures, Liouville measure to count states, quasiprobability, ...) and point to a straightforward generalization of such structures.\footnote{Note that we believe that the requirements we set, though necessary, are not sufficient to generalize the Lie algebra given by the classical Poisson brackets and the quantum commutators, meaning they cannot fully recover mechanics.
-Nonetheless, they are enough to recover measure theoretic structures, the topic of this paper.}
-
-The starting point is that any physical law deals with experimentally reproducible relationships.
-This implies a physical theory must give us the space of statistical ensembles (i.e.~the collection of all possible preparation procedures) that the theory allows.
-As a basic requirement, ensembles must be experimentally well-defined (leading to a topological space), allow statistical mixing (leading to a convex structure) and quantify the variability of the elements of the ensembles (leading to an entropy function).\footnote{The main point of departure from Generalized Probabilistic Theories (GPTs) is the entropy requirement.
-We found that a convex set is not enough to fully characterize an ensemble space, especially in the infinite dimensional case.} Additionally, though not necessary, we require having enough experimentally measurable quantities to identify each ensemble (i.e.~there is a family of affine functions that identify each ensemble through their values).
-Mathematically, these premises lead us to assume that an ensemble space is a convex subset of a metrizable second countable locally convex topological vector space.
-The question is, then, what generalized measure theoretic tool can quantify the number of states (reducing to the Liouville measure in the classical case) or the number of possible outcomes of a measurement (reducing to the dimension of a subspace in the quantum case)?
-What measure theoretic tool can represent each ensemble?
-If we are to define a physically meaningful notion of probability density, both of these questions must be answered.
-
-\textbf{Main finding.}  Choquet theory gives us a representation of ensembles in terms of convex combination of extreme points (i.e.~pure states).
-Classical spaces are those for which this representation is unique.
-Psuedo-probability measures are affine representations of the ensemble space, and different representation chosen based on the problem at hand.
-The probability measure finds a natural non-additive generalization in the fraction capacity $\frcap_{\ens}(A)$, which returns the highest possible fraction that an ensemble $\ens$ can be recovered from a convex combination of elements of $A$.
-The state counting measures finds a natural non-additive generalization in the state capacity $\capacity(A)$ which returns the highest exponential of the entropy (i.e.~entropic volume) achievable with a convex combination of elements of $A$.
+The default approach is simply to view quantum states as collections of probability measures over different quantities.
+Another approach is to extend probability measures to signed measures\cite{zachos2005,mandel1995,monchietti2025}, though it is not clear what negative probability represent, particularly since non-negative quasiprobability representations of quantum spaces exist.
+Yet another approach is to replace the $\sigma$-algebra with a different structure (CITE some quantum logic, quantum measure theory ).
+The situation, then, is that we multiple approaches, each with its own strength and weakness, with no clear connection or straight forward physical justification.
+Moreover, none of them generalize the connection between entropy and the classical Liouville measure for counting states.
+
+The aim of this paper is to present a simple, physically motivated structure in which all these approaches can be connected.
+Additionally, we show that a natural extension of probability and state counting measures exist under very general conditions.
+
+The main findings are the following:
+\begin{itemize}
+	\item any physical theory needs to provide a space of statistical ensemble, which need to be a bounded convex subsets of a locally convex second countable topological vector space
+	\item Choquet theory represents each ensemble $\ens$ with one or more probability measures $p_{\ens}$ on the set of extreme points $X$ (i.e. pure states)
+	\item if this representation is unique (i.e. the ensemble space is a simplex) classical probability theory can be used
+	\item the classical case corresponds to pure states being mutually exclusive
+	\item quasiprobability approaches provide affine representions of ensembles in terms of a function space, whose points are unrelated to the extreme points
+	\item the non-positivity of the measures is not key to quasiprobability approaches, and classical spaces admit quasiprobability as well
+	\item the algebras used in quantum logic and quantum measure theory can always be understood as sub-algebras of the Borel algebra of ensembles
+	\item a generalization of probability measures $p_{\ens}(A)$ emerges by asking the question ``what is the biggest fraction of a particular ensemble $\ens$ that can be expressed as a mixture of ensembles of $A$?''
+	\item on the set of extreme points $X$, the generalized probability measure is the supremum of all Choquet representations, and becomes additive in the classical case
+	\item a generalization of the state counting measure $\mu(A)$ emerges by asking ``what is the biggest spread that can be achieved with a mixture of ensembles of $A$?''
+	\item  the generalized state counting measure reduces to the Liouville measure in the classical case, and to the dimensionality of subspaces in the quantum case
+	\item to properly generalize probability theory, a non-additive notion of integral and derivative is needed which does not seem to match standard approaches.
+\end{itemize}
+
+A final remark before we start.
+Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects, such as information geometry, functional analysis, quantum logic and so on.
+At this point in time, what is needed is a synthesis of many different idea.
+However, such work carried to completion would be effectively overwhelming both to produce and to read.
+Therefore, the main part of the paper can only highlight the main conceptual results regarding measure theoretic representations.
+In the appendix wit go through some of the relevant mathematical detail while all the rest is left to our evolving collaborative open source work.\cite{aop-book}
 
@@
 
-A final remark before we start.
-Too often works in the foundations of physics focus on a particular aspect without concern to a broader connection to all relevant mathematical, physical and philosophical aspects, such as information geometry, functional analysis, quantum logic and so on.
-However, such a complete work would be effectively overwhelming both to produce and to read.
-The main part of the paper, then, highlights the main conceptual results regarding measure theoretic representations, the appendix goes through the relevant mathematical detail while all the rest is left to our evolving collaborative open source work.\cite{aop-book}
-
 \section{Ensemble space}
@@
 \textbf{Basic requirements for an ensemble space}.
-In our larger work, we strive to give all axioms and mathematical definitions solid grounding in physical requirements.\footnote{We call this approach Physical Mathematics, as starting from physical ideas recovers the mathematical structures, and only the mathematical structures, that are physically significant.} This way we can develop mathematical frameworks that are physically well founded.
-Specifically, we are developing a general theory of ensemble spaces, the full treatment of which is beyond the scope of this work.
-Here we limit ourselves to giving only the initial premises and the final mathematical characterization, with more details in the appendix and in our ongoing work.\cite{aop-book}
-
-We start by assuming that physical laws deal with experimentally reproducible relationships, which implicitly requires the notion of statistical ensembles.
-Conceptually, a statistical ensemble is the collection of outputs, taken at once, of a reproducible preparation procedure for a physical system.
+In our larger work, we strive to give all axioms and mathematical definitions solid grounding in physical requirements.\footnote{We call this approach Physical Mathematics, which stems from the observation that when physical objects are mapped to the right mathematical objects, the physical specification maps to the mathematical definitions.} Here we will only present a summary for the present investigation, leaving details to the appendix and our ongoing work.\cite{aop-book}
+
+Since physical laws deal with experimentally reproducible relationships, they are inherently relationships between statistical ensembles.
+Therefore, any physical theory must define the space of all ensembles allowed within the theory.
 In the classical case, this corresponds to a probability distribution over a discrete set of outcomes (e.g.
@@
 a positive semi-definite self-adjoint operator of trace one).
-A physical law, then, ultimately expresses relationships among ensembles.
-A minimal requirement for a physical theory, then, is to provide an ensemble space: the collection of all possible ensembles within that theory, together with a basic mathematical structure that all physical theories must possess.
-
-An ensemble space $\Ens$ will need to satisfy three basic requirements.
-First, it needs to guarantee experimental verifiability.
-Mathematically, the ensemble space must be a $\mathsf{T}_0$ second countable topological space, where each open set represents an experimentally verifiable statement\cite{aop-book,kelly1996}.\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Second, it needs to allow statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.
-Mathematically, this constrains the ensemble space to be a convex set.\footnote{Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i=1}^{\infty} p_i \ens_i$ converges in the space or not.} Third, an entropy function must be defined that characterizes the variability of the elements within an ensemble.
-
-Additionally, we will require that ensembles can be identified through real valued statistical quantities.\footnote{It is an open question whether this is an additional assumption or it can be derived from the other requirements.} A statistical quantity represents the expectation of a real valued quantity, and is therefore an affine continuous function $F : \Ens \to \mathbb{R}$.
+
+In this work, we will take an ensemble space $\Ens$ to be a bounded convex subset of a locally convex metrizable second countable topological vector space endowed with an entropy function.
+The topology captures experimental verifiability as open sets represent experimentally verifiable statements\cite{aop-book,kelly1996}.\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Convexity, and linearity more in general, comes from allowing statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.\footnote{Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i=1}^{\infty} p_i \ens_i$ converges in the space or not.} Lastly, the entropy characterizes the variability of the elements within an ensemble.
+In this context, a statistical quantity, representing the expectation of a real valued quantity, is an affine continuous function $F : \Ens \to \mathbb{R}$.
 That is, $F(p \ens[a] + \bar{p} \ens[b]) = p F(\ens[a]) + \bar{p} F(\ens[b])$.
-The last requirement is that there exists a family $F_i$ of statistical quantities that fully identify each ensemble.
-That is, $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i$.
 

### 2025-12-23 Articles/2024-NonAdditiveProbability/NonAdditiveProbability.tex
NonAdditiveProbability: more trimming

@@
 
-Since physical laws deal with experimentally reproducible relationships, they are inherently relationships between statistical ensembles.
-Therefore, any physical theory must define the space of all ensembles allowed within the theory.
-In the classical case, this corresponds to a probability distribution over a discrete set of outcomes (e.g.
-a fair coin) or over phase space (e.g.
-the symplectic manifold representing position and momentum for a classical particle).
-In the quantum case, this corresponds to a mixed state (e.g.
-a positive semi-definite self-adjoint operator of trace one).
-
-In this work, we will take an ensemble space $\Ens$ to be a bounded convex subset of a locally convex metrizable second countable topological vector space endowed with an entropy function.
-The topology captures experimental verifiability as open sets represent experimentally verifiable statements\cite{aop-book,kelly1996}.\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Convexity, and linearity more in general, comes from allowing statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.\footnote{Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i=1}^{\infty} p_i \ens_i$ converges in the space or not.} Lastly, the entropy characterizes the variability of the elements within an ensemble.
-In this context, a statistical quantity, representing the expectation of a real valued quantity, is an affine continuous function $F : \Ens \to \mathbb{R}$.
-That is, $F(p \ens[a] + \bar{p} \ens[b]) = p F(\ens[a]) + \bar{p} F(\ens[b])$.
-
-Each physical theory satisfies these axioms, even though it may be represented very differently mathematically.
-Classical mechanics, for example, is typically described in terms of the geometry of phase space.
-Quantum mechanics is most often defined in terms of complex inner product spaces, but can be reframed in terms of quasiprobability distributions.
-Therefore, while the above requirements are needed for a physical theory, it does not necessarily mean that it is convenient to describe problems or make calculations in terms of those definitions.
-In fact, those requirements interact with each other leading to a rich mathematical structure that connects to many of the structures we use in mathematical physics, and our goal is mainly to see how the different representations are connected.
-In particular, here we want to examine the different ways that measure theory can be used to represent ensembles in any given physical theory.
-
-% space is metrizable https://proofwiki.org/wiki/Birkhoff-Kakutani_Theorem/Topological_Vector_Space
-
-\textbf{Ensemble spaces as convex sets.} For the purpose of measure theoretic representations, it is best to understand an ensemble space $\Ens$ as a convex subset of a metrizable second countable locally convex topological vector space $V$.\footnote{To see the connection between the requirements and this mathematical characterization see the appendix.} See Fig.~\ref{convex_examples} for examples.
-The entropy is a function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
-\begin{equation}
-	\begin{aligned}
-		p S(\ens[a]) + \bar{p} &S(\ens[b]) \leq S(p \ens[a] + \bar{p} \ens[b]) \\
-		&\leq I(p,\bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])
-	\end{aligned}
-\end{equation}
-where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{The Shannon entropy is recovered as the only upper bound possible.}  Intuitively, the variability (i.e.~the entropy) cannot decrease during mixing, and will have an upper bound when mixing two ensembles whose instances are always distinguishable.
-
-Using the convex structure, we say that two ensembles are separate, noted $\ens[a] \separate \ens[b]$, if they cannot be expressed as a mixture of a common ensemble.
-That is, there is no $\ens[c]$ such that $\ens[a] = \alpha \ens[c] + \bar{\alpha} \ens[d]$ and $\ens[b] = \beta \ens[c] + \bar{\beta} \ens$ from some $\alpha, \beta \in (0,1]$ and $\ens[d],\ens \in \Ens$.
-No ensemble is separate from itself.
-An internal point is an ensemble that is not separate from any ensemble (i.e.~it is the mixture of any ensemble).
-An extreme point is an ensemble that is separate from all others (i.e.~it is only a mixture of itself).
-
 \begin{figure}[H]
@@
 	\end{tikzpicture}
-	\caption{Examples of a classical discrete ensemble space with three pure states (triangle) and of a two dimensional quantum ensemble space (ball).}\label{convex_examples}
+	\caption{Examples of a classical discrete ensemble space with three pure states (triangle) and of a two dimensional quantum ensemble space (ball).
+Infinite dimensional spaces, like continuous classical ensemble spaces and quantum ensemble space with continuous quantities cannot be visually represented.
+}\label{convex_examples}
 \end{figure}
 
-The set of all extreme points $X$, as we will see, plays a major role.
+Since physical laws deal with experimentally reproducible relationships, they are inherently relationships between statistical ensembles.
+Therefore, any physical theory must define the space of all ensembles allowed within the theory.
+In the classical case, this corresponds to a probability distribution over a discrete set of outcomes (e.g.
+a fair coin) or over phase space (e.g.
+the symplectic manifold representing position and momentum for a classical particle).
+In the quantum case, this corresponds to a mixed state (e.g.
+a positive semi-definite self-adjoint operator of trace one).
+See Fig.~\ref{convex_examples} for examples.
+
+In this work, we will take an ensemble space $\Ens$ to be a bounded convex subset of a locally convex metrizable second countable topological vector space endowed with an entropy function.
+The second countable topology captures experimental verifiability as open sets represent experimentally verifiable statements\cite{aop-book,kelly1996}.\footnote{A verifiable statement is an assertion for which there exists an experimental test that finishes successfully in finite time if and only if the statement is true.} Convexity, and linearity more in general, comes from allowing statistical mixtures: if $\ens[a],\ens[b] \in \Ens$ are ensembles, then $p \ens[a] + \bar{p} \ens[b]$ represents the statistical mixture where $\ens[a]$ is taken with probability $p$ and $\ens[b]$ with probability $\bar{p} = 1 - p$.\footnote{Only finite mixtures are guaranteed, while the topology will decide whether an infinite mixture $\sum_{i=1}^{\infty} p_i \ens_i$ converges in the space or not.} Lastly, the entropy characterizes the variability of the elements within an ensemble.
+The entropy is a function $S : \Ens \to \mathbb{R}$ that satisfies the following bounds
+\begin{equation}
+	\begin{aligned}
+		p S(\ens[a]) + \bar{p} &S(\ens[b]) \leq S(p \ens[a] + \bar{p} \ens[b]) \\
+		&\leq I(p,\bar{p}) + p S(\ens[a]) + \bar{p} S(\ens[b])
+	\end{aligned}
+\end{equation}
+where $I(p,\bar{p}) = - p \log p - \bar{p} \log(\bar{p})$.\footnote{The Shannon entropy is recovered as the only upper bound possible.}  Intuitively, the variability (i.e.~the entropy) cannot decrease during mixing, and will have an upper bound when mixing two ensembles whose instances are always distinguishable.
+
+In this context, a statistical quantity, representing the expectation of a real valued quantity, is an affine continuous function $F : \Ens \to \mathbb{R}$.
+That is, $F(p \ens[a] + \bar{p} \ens[b]) = p F(\ens[a]) + \bar{p} F(\ens[b])$.
+We also assume that there are enough statistical quantities to identify each ensemble.
+That is, there is a family $F_i$ such that $\ens[a] = \ens[b]$ if and only if $F_i(\ens[a]) = F_i(\ens[b])$ for all $i \in I$.
+This connects to the typical constructions for function spaces and their dual.
+
+Since any physical theory must provide experimental verifiability, statistical mixing and an entropy function, this gives us a general and solid conceptual foundation.
+While we are still in the process of understanding exactly how these requirements give us local convexity and metrizability, for the purpose of this work, we will simply assume them.
+
+% space is metrizable https://proofwiki.org/wiki/Birkhoff-Kakutani_Theorem/Topological_Vector_Space
+
+\textbf{Extreme points and compactness.} Using the convex structure, we say that two ensembles are separate, noted $\ens[a] \separate \ens[b]$, if they cannot be expressed as a mixture of a common ensemble.
+That is, there is no $\ens[c]$ such that $\ens[a] = \alpha \ens[c] + \bar{\alpha} \ens[d]$ and $\ens[b] = \beta \ens[c] + \bar{\beta} \ens$ from some $\alpha, \beta \in (0,1]$ and $\ens[d],\ens \in \Ens$.
+No ensemble is separate from itself.
+An internal point is an ensemble that is not separate from any ensemble (i.e.~it is the mixture of any ensemble).
+An extreme point is an ensemble that is separate from all others (i.e.~it is only a mixture of itself).
+
+The set of all extreme points $X$, as we will see, plays a major role in Choquet theory.
 However, on physical grounds, these may not be proper ensembles.
@@
 Additionally, we require the existence of enough statistical quantities to identify each ensemble.
-Under these assumptions, an ensemble space $\Ens$ is a convex set of a metrizable second countable locally convex topological vector space $V$.
+We therefore take an ensemble space $\Ens$ to be a bounded convex subset of a metrizable second countable locally convex topological vector space $V$.
 In some cases, for simplicity, we will require $\Ens$ to admit a ``compact closure'' $E$, so that it includes the set $X$ of its extreme points.
@@
 The expectation is linear with respect to the variable (i.e.~$E[aF+bG]= aE[F]+bE[G]$).
-In our case, what is more interesting is that it is an affine of the probability measure: if $p = \lambda p_1 + \bar{\lambda} p_2$ then $\int_X F dp = \lambda \int_X F dp_1 + \bar{\lambda} \int_X F dp_2$.
+In our case, what is more interesting is that it is an affine function of the probability measure: if $p = \lambda p_1 + \bar{\lambda} p_2$ then $\int_X F dp = \lambda \int_X F dp_1 + \bar{\lambda} \int_X F dp_2$.
 That is, as we required for statistical variables, the expectation for a mixture is the mixture of the expectations.
@@
 This confirms the physical intuition from classical and quantum statistical mechanics.
+
+TODO add picture
 
